### 2016-04-27 capability-machine.tex
Initial document for formalisation of capability machine.

\documentclass{article}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{thmtools}
\usepackage{stmaryrd}

\declaretheorem[numbered=yes,name=Lemma]{lemma}
\declaretheorem[numbered=yes,name=Definition]{definition}

\newcommand{\update}[2]{[#1 \mapsto #2]}%TODO change this to nicer notation \Phi[heap.a \mapsto w]

\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hv}{hv}
\newcommand{\rv}{rv}
\newcommand{\lv}{lv}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{a}
\newcommand{\word}{w}
\newcommand{\len}{len}
\newcommand{\mem}{mem}
\newcommand{\reg}{reg}
\newcommand{\heap}{heap}
\newcommand{\mode}{mode}
\newcommand{\perm}{perm}
\newcommand{\failed}{failed}
\newcommand{\halted}{halted}
\newcommand{\false}{\mathit{false}}
\newcommand{\true}{\mathit{true}}
\newcommand{\decode}{\mathit{decode}}
\newcommand{\updatePcPerm}{\mathit{updatePcPerm}}
\newcommand{\executeAllowed}{\mathit{executeAllowed}}
\newcommand{\nonZero}[1]{\mathit{nonZero}(#1)}
\newcommand{\readAllowed}[1]{\mathit{readAllowed}(#1)}
\newcommand{\writeAllowed}[1]{\mathit{writeAllowed}(#1)}

\newcommand{\plaindom}[1]{\mathrm{#1}}
\newcommand{\Caps}{\plaindom{Cap}}
\newcommand{\Words}{\plaindom{Word}}
\newcommand{\Addrs}{\plaindom{Addr}}
\newcommand{\Mems}{\plaindom{Mem}}
\newcommand{\RegName}{\plaindom{RegName}}
\newcommand{\Regs}{\plaindom{Reg}}
\newcommand{\Heaps}{\plaindom{Heap}}
\newcommand{\Confs}{\plaindom{Conf}}
\newcommand{\Lens}{\plaindom{Len}}
\newcommand{\Instrs}{\plaindom{Instructions}}
\newcommand{\nats}{\mathbb{N}}

\newcommand{\Perms}{\plaindom{Perm}}

\newcommand{\refreg}[1]{\lfloor #1 \rfloor}
\newcommand{\refheap}[1]{\langle #1 \rangle_h}

\newcommand{\instr}[1]{\mathtt{#1}}
\newcommand{\fail}{\instr{fail}}
\newcommand{\halt}{\instr{halt}}
\newcommand{\oneinstr}[2]{\instr{#1} \; #2}
\newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}

\newcommand{\twoinstr}[3]{\instr{#1} \; #2 \; #3}
\newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
\newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
\newcommand{\setptr}[2]{\twoinstr{setptr}{#1}{#2}}
\newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
\newcommand{\store}[2]{\twoinstr{store}{#1}{#2}}
\newcommand{\load}[2]{\twoinstr{load}{#1}{#2}}
\newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}

\newcommand{\threeinstr}[4]{\instr{#1} \; #2 \; #3 \; #4}
\newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
\newcommand{\subseg}[3]{\threeinstr{subseg}{#1}{#2}{#3}}
\newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}

\newcommand{\plainperm}[1]{\mathrm{#1}}
\newcommand{\readonly}{\plainperm{ro}}
\newcommand{\readwrite}{\plainperm{rw}}
\newcommand{\exec}[1]{\plainperm{x_{#1}}}
\newcommand{\entry}[1]{\plainperm{e_{#1}}}
\newcommand{\key}{\plainperm{k}}

\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}

\begin{document}
$\RegName$ contains $\pcreg$, but is otherwise some undefined, finite set.
\begin{align*}
\Addrs &::= \nats & & &
\Words &::= \Caps + \nats \\
\Regs  &::= \RegName \rightarrow \Words & & &
\Heaps &::= \Addrs \rightarrow \Words \\
\Perms &::= \{\readonly, \readwrite, \exec{u}, \exec{p}, \entry{u}, \entry {p}, \key\} & & &
\Mems  &::= \Regs \times \Heaps \\
\Lens  &::= \nats & & &
\Caps  &::= \Perms \times \Lens \times \Addrs \\
\Confs &::= \Mems + \{\failed, \halted\}
\end{align*}
Notation:
$$\begin{array}{rcl}
i      &\in& \Instrs \\
r      &\in& \RegName\\
\mem   &::=& (\reg,\heap)\\
\len   &\in& \nats \\
\pc    &\in& \Caps \\
\pcreg &\in& \RegName \\
\Phi   &::=& \mem \in \Confs\\
\addr      &\in& \Addrs\\
\perm  &\in& \Perms\\
(\perm,len,\addr) &\in& \Caps \\
\end{array}$$
Further definitions:
$$\begin{array}{rcl}
\lv    &::=& \refreg{r} \\
\hv    &::=& \refheap{r}\\
\rv    &::=& n \mid \lv \\
i      &::=& \fail \mid \halt \mid 
             \jmp{\lv} \mid \jnz{\lv}{\rv} \mid
             \isptr{\lv}{\rv} \mid \setptr{\lv}{\rv} \mid \\
       &   & \lea{\lv}{\rv} \mid\move{\lv}{\rv} \mid \load{\lv}{\hv} \mid \store{\hv}{\rv} \mid  \\
       &   & \restrict{\lv}{\rv}{\rv} \mid \subseg{\lv}{\rv}{\rv} \mid \plus{\lv}{\rv}{\rv}
\end{array}$$

\begin{align*}
\decode &:\nats \rightarrow \Instrs
\end{align*}
%TODO for some decode function

\begin{align*}
\Phi & \rightarrow \sem{\decode(\Phi.\reg(\pcreg))}(\Phi) & & \text{if $\Phi.\reg(\pcreg) = (\exec{\_},{\_},{\_})$} \\
\Phi & \rightarrow \failed                                      & & \text{otherwise}
\end{align*}
\begin{align*}
  \executeAllowed(\perm) &=
                           \begin{cases}
                             \true & \text{if } \perm \in \{\exec{u},\exec{p},\entry{u},\entry{p}\} \\
                             \false & \text{otherwise}
                           \end{cases} \\
  \readAllowed{\perm} &=
                           \begin{cases}
                             \true & \text{if } \perm \in \{\readonly, \readwrite, \exec{u},\exec{p}\} \\
                             \false & \text{otherwise}
                           \end{cases} \\
  \writeAllowed{\perm} &=
                           \begin{cases}
                             \true & \text{if } \perm \in \{\readwrite\} \\
                             \false & \text{otherwise}
                           \end{cases} \\
  \updatePcPerm (\perm,\len,\addr) &=
                                     \begin{cases}
                                       (\perm,\len,\addr) & \text{if $\perm\in\{\exec{u},\exec{p}\}$} \\
                                       (\exec{\var{m}},\len,\addr) & \text{if $\perm = \entry{\var{m}}$}
                                     \end{cases} \\
  \nonZero{w} &=
                \begin{cases}
                  \true & \text{if $w\in \Caps$ or $w\in \nats$ and $w \neq 0$}\\
                  \false & \text{otherwise}
                \end{cases}
\end{align*}
%\Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
\begin{align*}
  \sem{\fail}(\Phi)     & = \failed \\
  \sem{\halt}(\Phi)     & = \halted \\
  \sem{\jmp{\lv}}(\Phi) & = \begin{cases}
                            (\Phi.\reg\update{\pcreg}{\updatePcPerm(c)}) & \text{if }\Phi.reg(lv) = c \\
                                                                         & \text{  and }c=(\perm,\len,\addr)\\
                                                                         & \text{  and }\executeAllowed(\perm)\\
                            \failed & \text{otherwise }
                            \end{cases} \\
  \sem{\jnz{\lv}{\rv}}(\Phi) & = \begin{cases}
                            \Phi.\reg\update{\pcreg}{\updatePcPerm(\var{c})} &
                            \begin{array}{l}
                              \text{if $\nonZero{\Phi.\reg(\rv)}$} \\ 
                              \text{  and $\Phi.reg(lv) = c$} \\
                              \text{  and $c=(\perm,\len,\addr)$}\\
                              \text{  and $\executeAllowed(\perm)$}
                            \end{array}
                            \\ %TODO Maybe combine with jump. (failed + this)
                            \Phi.\reg\update{\pcreg}{\Phi.\reg(\pcreg) + 1} & \text{if not $\nonZero{\Phi.\reg(\rv)}$}\\
                            \failed & \text{otherwise }
                            \end{cases} \\
 \sem{\load{\refreg{r_1}}{\refheap{r_2}}} & =
                                 \begin{cases}
                                   \Phi.\reg\update{r_1}{\var{w}} &
                                   \begin{array}{l}
                                     \text{if }\Phi.\reg(r_2) = (\perm,\len,\addr)\\
                                     \text{  and }\readAllowed{\perm} \text{ and } \var{w} = \Phi.\heap(\addr)
                                   \end{array}\\
                                   \failed & \text{otherwise }
                                 \end{cases}\\
 \sem{\store{\refheap{r_1}}{\refreg{r_2}}} & =
                                 \begin{cases}
                                   \Phi.\heap\update{\addr}{\var{w}} &
                                   \begin{array}{l}
                                     \text{if }\Phi.\reg(r_1) = (\perm,\len,\addr)\\
                                     \text{  and }\writeAllowed{\perm} \text{ and } \var{w} = \Phi.\reg(r_2)
                                   \end{array}\\
                                   \failed & \text{otherwise }
                                 \end{cases}\\
 \sem{\move{\refreg{r_1}}{\rv}} & =
                                 \begin{cases}
                                   \Phi.\reg\update{r_1}{\Phi.\reg(\rv)} & \text{if $r_1 \neq \pcreg$} \\
                                   \failed   & \text{otherwise }
                                 \end{cases}\\
\end{align*}

\end{document}

### 2016-04-29 capability-machine.tex
Add a list of related papers and documents.

@@
 \usepackage{stmaryrd}
+\usepackage{natbib}
+\usepackage{url}
 
@@
 
+\section{Related reading}
+\label{sec:related-reading}
+
+This is a list of related work that might be interesting to read in the context
+of this project.
+
+\subsection{Capability machines}
+\label{sec:rw-cap-machines}
+
+\subsubsection{M-Machine}
+More than 20 years ago, \cite{Carter:1994:HSF:195473.195579} have described the
+use of capabilities in the M-Machine. They do seem to have a reference for the
+instruction set after all~\citep{Dally1997Memo59}; it seems like the server was
+just temporarily down when we were looking for this the first time...
+
+\subsubsection{CHERI}
+
+The CHERI processor is a much more recent capability machine, described
+by~\cite{Woodruff:2014:CCM:2665671.2665740,Watson2015Cheri}.
+
+Another result of this project is also CheriBSD: an adaptation of FreeBSD to the
+CHERI
+processor.\footnote{\url{http://www.cl.cam.ac.uk/research/security/ctsrd/cheri/cheribsd.html}}
+It is not separately described in a published paper, but mentioned in the papers
+cited above and in some tech reports (see url). This work includes a
+pure-capability ABI that could provide some interesting examples.
+
+The CHERI team also has a webpage with all of their CHERI-related publications
+(including TRs and
+such)\footnote{\url{http://www.cl.cam.ac.uk/research/security/ctsrd/cheri/}}.
+
+\subsection{Logical Relations}
+\label{sec:rw-log-rel}
+
+Some papers on logical relations that are relevant for this work are the
+following:
+
+\cite{Hur:2011:KLR:1926385.1926402} describe a logical relation between ML and
+a (standard) assembly language for expressing compiler correctness.  Relevant
+because they target an assembly language, and they use biorthogonality.
+
+\cite{Dreyer:2010:IHS:1863543.1863566} describe a logical relation for a ML-like
+language and use public/private transitions to reason about well-bracketed
+control flow. Relevant because we are considering to cover an example of
+enforcing well-bracketed control flow in a capability machine.
+
+\cite{Devriese:2016ObjCap} describe a logical relation for a JavaScript-like
+language with object capabilities.  Relevant because it treats object
+capabilities, albeit in a JavaScript-like lambda calculus.
+
+\bibliographystyle{plainnat}
+\bibliography{refs}
+
 \end{document}

### 2016-05-02 capability-machine.tex
pictures from Dominique's visit
Wrote up what we talked about during Dominique's visit

@@
 
+\usepackage{tikz}
+
 \declaretheorem[numbered=yes,name=Lemma]{lemma}
@@
 \newcommand{\update}[2]{[#1 \mapsto #2]}%TODO change this to nicer notation \Phi[heap.a \mapsto w]
+\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
 
+%Math notation
+\newcommand{\parfun}{\rightharpoonup}
+\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
+\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
+\newcommand{\union}{\mathbin{\cup}}
+\DeclareMathOperator{\dom}{dom}
+
+\newcommand{\false}{\mathit{false}}
+\newcommand{\true}{\mathit{true}}
+
+%Variables
 \newcommand{\var}[1]{\mathit{#1}}
-\newcommand{\hv}{hv}
-\newcommand{\rv}{rv}
-\newcommand{\lv}{lv}
+\newcommand{\hv}{\var{hv}}
+\newcommand{\rv}{\var{rv}}
+\newcommand{\lv}{\var{lv}}
 \newcommand{\pc}{\mathit{pc}}
 \newcommand{\pcreg}{\mathrm{pc}}
-\newcommand{\addr}{a}
-\newcommand{\word}{w}
-\newcommand{\len}{len}
-\newcommand{\mem}{mem}
-\newcommand{\reg}{reg}
-\newcommand{\heap}{heap}
-\newcommand{\mode}{mode}
-\newcommand{\perm}{perm}
-\newcommand{\failed}{failed}
-\newcommand{\halted}{halted}
-\newcommand{\false}{\mathit{false}}
-\newcommand{\true}{\mathit{true}}
-\newcommand{\decode}{\mathit{decode}}
-\newcommand{\updatePcPerm}{\mathit{updatePcPerm}}
-\newcommand{\executeAllowed}{\mathit{executeAllowed}}
-\newcommand{\nonZero}[1]{\mathit{nonZero}(#1)}
-\newcommand{\readAllowed}[1]{\mathit{readAllowed}(#1)}
-\newcommand{\writeAllowed}[1]{\mathit{writeAllowed}(#1)}
+\newcommand{\addr}{\var{a}}
+\newcommand{\offset}{\var{offset}}
+\newcommand{\word}{\var{w}}
+\newcommand{\start}{\var{base}}
+\newcommand{\addrend}{\var{end}}
+\newcommand{\mem}{\var{mem}}
+\newcommand{\reg}{\var{reg}}
+\newcommand{\heapseg}{\var{hs}}
+\newcommand{\heap}{\var{heap}}
+\newcommand{\mode}{\var{mode}}
+\newcommand{\perm}{\var{perm}}
 
+\newcommand{\stdcap}[1][\perm]{\left(#1,\start,\addrend,\offset \right)}
+
+%Configuration end states
+\newcommand{\failed}{\mathit{failed}}
+\newcommand{\halted}{\mathit{halted}}
+
+%Functions
+\newcommand{\plainfun}[1]{\mathit{#1}}
+\newcommand{\decode}{\plainfun{decode}}
+\newcommand{\encode}{\plainfun{encode}}
+\newcommand{\encodePerm}{\plainfun{encodePerm}}
+\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}}
+\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
+\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
+\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
+\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
+
+%World operations
+\newcommand{\future}{\mathbin{\sqsupseteq}}
+\newcommand{\heapSat}[3][\Phi]{#1 :_{#2} #3}
+
+%Assembly labels
+\newcommand{\codelabel}[1]{\mathit{#1}}
+\newcommand{\init}{\codelabel{init}}
+\newcommand{\malloc}{\codelabel{malloc}}
+
+%Type(s)
+\newcommand{\type}[1]{\mathrm{#1}}
+\newcommand{\asmType}{\plaindom{AsmType}}
+
+%Domains
 \newcommand{\plaindom}[1]{\mathrm{#1}}
@@
 \newcommand{\Mems}{\plaindom{Mem}}
-\newcommand{\RegName}{\plaindom{RegName}}
+\newcommand{\RegName}{\plaindom{RegisterName}}
 \newcommand{\Regs}{\plaindom{Reg}}
@@
 \newcommand{\Confs}{\plaindom{Conf}}
-\newcommand{\Lens}{\plaindom{Len}}
 \newcommand{\Instrs}{\plaindom{Instructions}}
 \newcommand{\nats}{\mathbb{N}}
-
+\newcommand{\ints}{\mathbb{Z}}
 \newcommand{\Perms}{\plaindom{Perm}}
 
+\newcommand{\RegionName}{\plaindom{RegionName}}
+\newcommand{\Worlds}{\plaindom{World}}
+\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
+
+%LR
+\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
+\newcommand{\valueintr}[1]{\intr{V}{#1}}
+\newcommand{\exprintr}[1]{\intr{E}{#1}}
+\newcommand{\contintr}[1]{\intr{K}{#1}}
+\newcommand{\stdvr}{\valueintr{\asmType}}
+\newcommand{\stder}{\exprintr{\asmType}}
+\newcommand{\stdkr}{\contintr{\asmType}}
+\newcommand{\npair}[2][n]{\left(#1,#2 \right)}
+
+%Reference register/heap
 \newcommand{\refreg}[1]{\lfloor #1 \rfloor}
@@
 
-\newcommand{\instr}[1]{\mathtt{#1}}
+%Instructions
+%No arguments
 \newcommand{\fail}{\instr{fail}}
 \newcommand{\halt}{\instr{halt}}
+%One argument
+\newcommand{\instr}[1]{\mathtt{#1}}
 \newcommand{\oneinstr}[2]{\instr{#1} \; #2}
 \newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
-
+%Two arguments
 \newcommand{\twoinstr}[3]{\instr{#1} \; #2 \; #3}
@@
 \newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
-
+%Three arguments
 \newcommand{\threeinstr}[4]{\instr{#1} \; #2 \; #3 \; #4}
@@
 
+%Permissions
 \newcommand{\plainperm}[1]{\mathrm{#1}}
+\newcommand{\noperm}{\plainperm{o}}
 \newcommand{\readonly}{\plainperm{ro}}
 \newcommand{\readwrite}{\plainperm{rw}}
-\newcommand{\exec}[1]{\plainperm{x_{#1}}}
-\newcommand{\entry}[1]{\plainperm{e_{#1}}}
-\newcommand{\key}{\plainperm{k}}
+\newcommand{\exec}{\plainperm{rx}}
+\newcommand{\entry}{\plainperm{e}}
+\newcommand{\rwx}{\plainperm{rwx}}
 
-\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
+%OP sem
+\newcommand{\diverge}[1][n]{\not\Downarrow_{#1}}
 
@@
 \Addrs &::= \nats & & &
-\Words &::= \Caps + \nats \\
+\Words &::= \Caps + \ints \\
 \Regs  &::= \RegName \rightarrow \Words & & &
 \Heaps &::= \Addrs \rightarrow \Words \\
-\Perms &::= \{\readonly, \readwrite, \exec{u}, \exec{p}, \entry{u}, \entry {p}, \key\} & & &
+\Perms &::= \{ \noperm, \readonly, \readwrite, \exec, \entry, \rwx\} & & &
 \Mems  &::= \Regs \times \Heaps \\
-\Lens  &::= \nats & & &
-\Caps  &::= \Perms \times \Lens \times \Addrs \\
+\Caps  &::= \Perms \times \Addrs \times \Addrs \times \Addrs & & &
 \Confs &::= \Mems + \{\failed, \halted\}
 \end{align*}
+
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+  \node[main node] (1) {$\rwx$};
+  \node[main node] (2) [below right of=1] {$\exec$};
+  \node[main node] (3) [below of=2] {$\entry$};
+  \node[main node] (4) [below left of=1] {$\readwrite$};
+  \node[main node] (5) [below of=4] {$\readonly$};
+  \node[main node] (6) [below right of=5] {$\noperm$};
+
+  \path[every node/.style={font=\sffamily\small}]
+    (1) edge (2)
+    (2) edge (3)
+    (3) edge (6)
+    (1) edge (4)
+    (4) edge (5)
+    (5) edge (6);
+\end{tikzpicture}
+\caption{Permission hierarchy}
+\label{fig:perm-hier}
+\end{figure}
 Notation:
@@
 \mem    &::=& (\reg,\heap)\\
-\len   &\in& \nats \\
 \pc     &\in& \Caps \\
@@
 \perm   &\in& \Perms\\
-(\perm,len,\addr) &\in& \Caps \\
+(\perm,\start,\addrend,\addr) &\in& \Caps \\
 \end{array}$$
@@
 
+\subsection*{Semantics}
 \begin{align*}
-\decode &:\nats \rightarrow \Instrs
+\decode &:\ints \rightarrow \Instrs
 \end{align*}
@@
 \begin{align*}
-  \executeAllowed(\perm) &=
+  \executeAllowed{\perm} &=
                            \begin{cases}
-                             \true & \text{if } \perm \in \{\exec{u},\exec{p},\entry{u},\entry{p}\} \\
+                             \true & \text{if } \perm \in \{ \rwx, \exec, \entry \} \\
                              \false & \text{otherwise}
@@
                            \begin{cases}
-                             \true & \text{if } \perm \in \{\readonly, \readwrite, \exec{u},\exec{p}\} \\
+                             \true & \text{if } \perm \in \{ \rwx, \exec, \readwrite, \readonly \} \\
                              \false & \text{otherwise}
@@
                            \begin{cases}
-                             \true & \text{if } \perm \in \{\readwrite\} \\
+                             \true & \text{if } \perm \in \{ \rwx, \readwrite\} \\
                              \false & \text{otherwise}
                            \end{cases} \\
-  \updatePcPerm (\perm,\len,\addr) &=
+  \updatePcPerm{\perm,\start,\addrend,\addr} &=
                                      \begin{cases}
-                                       (\perm,\len,\addr) & \text{if $\perm\in\{\exec{u},\exec{p}\}$} \\
-                                       (\exec{\var{m}},\len,\addr) & \text{if $\perm = \entry{\var{m}}$}
+                                       (\perm,\start,\addrend,\addr) & \text{if $\perm\in\{ \rwx, \exec \}$} \\
+                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}
                                      \end{cases} \\
@@
                 \begin{cases}
-                  \true & \text{if $w\in \Caps$ or $w\in \nats$ and $w \neq 0$}\\
+                  \true & \text{if $w\in \Caps$ or $w\in \ints$ and $w \neq 0$}\\
                   \false & \text{otherwise}
@@
 \end{align*}
-%\Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
+%TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
+%TODO: Update semantics according to changes.
 \begin{align*}
@@
   \sem{\jmp{\lv}}(\Phi) & = \begin{cases}
-                            (\Phi.\reg\update{\pcreg}{\updatePcPerm(c)}) & \text{if }\Phi.reg(lv) = c \\
-                                                                         & \text{  and }c=(\perm,\len,\addr)\\
-                                                                         & \text{  and }\executeAllowed(\perm)\\
+                            (\Phi.\reg\update{\pcreg}{\updatePcPerm{c}}) & \text{if }\Phi.reg(lv) = c \\
+                                                                         & \text{  and }c=(\perm,\start,addrend,\addr)\\
+                                                                         & \text{  and }\executeAllowed{\perm}\\
                             \failed & \text{otherwise }
@@
   \sem{\jnz{\lv}{\rv}}(\Phi) & = \begin{cases}
-                            \Phi.\reg\update{\pcreg}{\updatePcPerm(\var{c})} &
+                            \Phi.\reg\update{\pcreg}{\updatePcPerm{\var{c}}} &
                             \begin{array}{l}
@@
                               \text{  and $\Phi.reg(lv) = c$} \\
-                              \text{  and $c=(\perm,\len,\addr)$}\\
-                              \text{  and $\executeAllowed(\perm)$}
+                              \text{  and $c=(\perm,\start,\addrend,\addr)$}\\
+                              \text{  and $\executeAllowed{\perm}$}
                             \end{array}
@@
                                    \begin{array}{l}
-                                     \text{if }\Phi.\reg(r_2) = (\perm,\len,\addr)\\
+                                     \text{if }\Phi.\reg(r_2) = (\perm,\start,\addrend,\addr)\\
                                      \text{  and }\readAllowed{\perm} \text{ and } \var{w} = \Phi.\heap(\addr)
@@
                                    \begin{array}{l}
-                                     \text{if }\Phi.\reg(r_1) = (\perm,\len,\addr)\\
+                                     \text{if }\Phi.\reg(r_1) = (\perm,\start,\addrend,\addr)\\
                                      \text{  and }\writeAllowed{\perm} \text{ and } \var{w} = \Phi.\reg(r_2)
@@
 \end{align*}
+\section{Examples}
+\label{sec:examples}
+\subsection{Ticket Dispenser}
+\newcommand{\size}{\var{size}}
+\label{sec:tick-disp}
+\[
+  \begin{array}{r l}
+    \init: & \move{\refreg{r_1}}{\size} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
+           & \lea{\refreg{r_0}}{3} \\
+           & \jmp{\malloc} \\
+           & \store{\refheap{r_1}}{(\encode(i_1))} \\
+           & \lea{\refreg{r_1}}{1} \\
+           & \store{\refheap{r_1}}{(\encode(i_2))} \\
+           & \lea{\refreg{r_1}}{1} \\
+           & \vdots \\
+           & \store{\refheap{r_1}}{(\encode(i_6))} \\
+           & \lea{\refreg{r_1}}{1} \\
+           & \move{\refreg{r_0}}{\refreg{r_1}} \\
+           & \lea{\refreg{r_0}}{1} \\
+           & \store{\refheap{r_1}}{\refreg{r_0}} \\
+           & \lea{\refreg{r_1}}{1} \\
+           & \store{\refheap{r_1}}{0} \\
+           & \lea{\refreg{r_1}}{-7} \\
+           & \restrict{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
+           & \lea{\refreg{r_0}}{4} \\
+           & \restrict{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
+           & \jmp{\codelabel{adv}}
+  \end{array}
+\]
+Here $\size$ is 8. $i_1 \dots i_7$ refers to the instructions in the following program:
+\[
+  \begin{array}{r l}
+    i_1 :& \load{\refreg{r_2}}{\size-1} \\
+    i_2 :& \load{\refreg{r_1}}{\refheap{r_2}} \\
+    i_3 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
+    i_4 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
+    i_5 :& \move{\refreg{r_2}}{0} \\
+    i_6 :& \jmp{\refreg{r_0}}
+  \end{array}
+\]
+The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{ | c | l | }
+    \hline
+    Addr. & Content \\ \hline
+    1   &  $i_1$ \\ \hline
+    2   &  $i_2$ \\ \hline
+    3   &  $i_3$ \\ \hline
+    4   &  $i_4$ \\ \hline
+    5   &  $i_5$ \\ \hline
+    6   &  $i_6$ \\ \hline
+    7   & Cap. for addr. 8 \\ \hline
+    8   & The counter \\ \hline  
+  \end{tabular}
+  \caption{The heap layout for the ticket dispenser.}
+  \label{tab:tick-disp-heap}
+\end{figure}
 
+\section{Logical Relation}
+\label{sec:logical-relation}
+\subsection{Worlds}
+\[
+\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monfun \UPred{\Addrs \parfun \Worlds}))
+\]
+
+\[
+\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
+\]
+\begin{align*}
+  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
+\union \\
+                   & \{ \npair{\stdcap[\noperm] } \} 
+\union \\
+                   & \{ \npair{\stdcap[readonly]} \mid \\
+                   & \quad \exists r \in \RegionName \ldotp \\
+                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
+\union \\
+                   & \{ \npair{\readwrite} \mid \\
+                   & \quad \exists r \in \RegionName \ldotp \\
+                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
+                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
+                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
+                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
+                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\Phi\update{a}{v}]{n''}{W''}
+\union \\
+                   & \{ \npair{\stdcap[\exec] = c} \mid \\
+                   & \quad (\exists r \in \RegionName \ldotp \\
+                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
+                   & \quad \land \\
+                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
+                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
+                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+\}
+\end{align*}
+\[
+\stdkr \; : \;  \Worlds \monfun \UPred{\Words}%TODO: What should this be?
+\]
+\begin{align*}
+  \stdkr(W) \defeq & \{ \npair{c} \mid \\
+                   & \quad \forall v\ldotp \in \stdvr(W) \Rightarrow \\
+                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
+                     \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
+                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \diverge \}
+\end{align*}
+
+\[
+\stder \; : \;  \Worlds \monfun \UPred{\Words}%TODO: What should this be?
+\]
+\begin{align*}
+  \stder(W) \defeq & \{ \npair{\reg} \mid \\
+                   & \quad \forall \npair{c} \in \stdkr(W) \\
+                   & \qquad  \forall \heapSat{\heap}{n}{W} \ldotp \\
+                   & \qquad \quad (\reg\update{r_0}{c},\heap) \diverge
+\end{align*}
+
+\begin{definition}[Heap staisfaction/erasure]
+\begin{align*}
+  \heapSat{n}{W} & & \text{iff} & &
+                                \begin{gathered}
+\exists \heapseg_{rem} \ldotp \Phi.\heap = \heapseg_{rem} \uplus \biguplus_{r\in \dom(W)} \heapseg_r\\
+\text{and}\\
+\forall r \in \dom(W) \ldotp \npair{\heapseg_r} \in W(r)(W)
+\end{gathered}
+\end{align*}
+\end{definition}
 \section{Related reading}

### 2016-05-09 capability-machine.tex
Small latex fix.
Added O relation and fixed a couple of typos.
still work in progress
Updated the ticket dispenser code to use a capability stored on the heap
Update some notations wrt updating heap/registers
Updated semantics and added the remaining instructions needed for example.

@@
 
-\newcommand{\update}[2]{[#1 \mapsto #2]}%TODO change this to nicer notation \Phi[heap.a \mapsto w]
+\newcommand{\update}[2]{[#1 \mapsto #2]}
 \newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
@@
 \DeclareMathOperator{\dom}{dom}
+
+\newcommand{\undefined}{\mathit{undefined}}
 
@@
 
+%Memory projections
+\newcommand{\plainproj}[1]{\mathrm{#1}}
+\newcommand{\memheap}[1][\Phi]{#1.\plainproj{heap}}
+\newcommand{\memreg}[1][\Phi]{#1.\plainproj{reg}}
+
+\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{heap}.#2}{#3}}
+\newcommand{\updateReg}[3][\Phi]{#1\update{\plainproj{reg}.#2}{#3}}
+
 %Configuration end states
-\newcommand{\failed}{\mathit{failed}}
-\newcommand{\halted}{\mathit{halted}}
+\newcommand{\failed}{\textsl{failed}}
+\newcommand{\halted}{\textsl{halted}}
 
@@
 \newcommand{\encodePerm}{\plainfun{encodePerm}}
-\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}}
+\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
 \newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
@@
 \newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
+\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}
 
@@
 \newcommand{\future}{\mathbin{\sqsupseteq}}
-\newcommand{\heapSat}[3][\Phi]{#1 :_{#2} #3}
+\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
 
@@
 \newcommand{\stdkr}{\contintr{\asmType}}
+\newcommand{\observations}{\mathcal{O}}
 \newcommand{\npair}[2][n]{\left(#1,#2 \right)}
@@
 \Caps  &::= \Perms \times \Addrs \times \Addrs \times \Addrs & & &
-\Confs &::= \Mems + \{\failed, \halted\}
+\Confs &::= \Mems + \{\failed, \halted \times \Mems\}
 \end{align*}
@@
 \Phi    &::=& \mem \in \Confs\\
+\memheap&\in& \Heaps \\
+\memreg &\in& \Regs \\
 \addr   &\in& \Addrs\\
@@
 (\perm,\start,\addrend,\addr) &\in& \Caps \\
+n       &\in& \ints\\
 \end{array}$$
@@
 \subsection*{Semantics}
+Assume a $\decode$ function that decodes integer to instructions:
 \begin{align*}
@@
 \end{align*}
-%TODO for some decode function
-
+Assume an $\encodePerm$ function that encodes a permission as an integer:
 \begin{align*}
-\Phi & \rightarrow \sem{\decode(\Phi.\reg(\pcreg))}(\Phi) & & \text{if $\Phi.\reg(\pcreg) = (\exec{\_},{\_},{\_})$} \\
+\encodePerm &: \Perms \rightarrow \ints
+\end{align*}
+\begin{align*}
+  \Phi & \rightarrow \sem{\decode(\memreg(\pcreg))}(\Phi) & &                                   
+                                                              \arraycolsep=0pt
+                                                              \begin{array}{l}
+                                                                \text{if $\memreg(\pcreg) = \stdcap$}\\
+                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
+                                                                \quad\text{and $\perm \in \{ \exec,\rwx \}$ }
+                                                              \end{array}\\
 \Phi & \rightarrow \failed                                 & & \text{otherwise}
@@
                   \false & \text{otherwise}
+                \end{cases} \\
+  \withinBounds{(\_,\start,\addrend,\addr)} &=
+                                              \begin{cases}
+                                                \true  & \text{if $\start \leq \addr \leq \addrend$} \\
+                                                \false & \text{otherwise}
                                               \end{cases}
@@
 %TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
-%TODO: Update semantics according to changes.
 \begin{align*}
   \sem{\fail}(\Phi)                        & = \failed \\
-  \sem{\halt}(\Phi)     & = \halted \\
-  \sem{\jmp{\lv}}(\Phi) & = \begin{cases}
-                            (\Phi.\reg\update{\pcreg}{\updatePcPerm{c}}) & \text{if }\Phi.reg(lv) = c \\
-                                                                         & \text{  and }c=(\perm,\start,addrend,\addr)\\
-                                                                         & \text{  and }\executeAllowed{\perm}\\
+  \sem{\halt}(\Phi)                        & = (\halted,\Phi) \\
+  \sem{\jmp{\lv}}(\Phi)                    & = 
+                                             \begin{cases}
+                                               \updateReg{\pcreg}{\updatePcPerm{c}} &
+                                               \arraycolsep=0pt
+                                               \begin{array}{l}
+                                                 \text{if }\memreg(lv) = c \\
+                                                 \quad\text{and }c=(\perm,\start,\addrend,\addr)\\
+                                                 \quad\text{and }\executeAllowed{\perm}
+                                               \end{array} \\
                                                \failed                              & \text{otherwise }
                                              \end{cases} \\
-  \sem{\jnz{\lv}{\rv}}(\Phi) & = \begin{cases}
-                            \Phi.\reg\update{\pcreg}{\updatePcPerm{\var{c}}} &
+  \sem{\jnz{\lv}{\rv}}(\Phi)               & = 
+                                             \begin{cases}
+                                               \updateReg{\pcreg}{\updatePcPerm{\var{c}}} &
+                                               \arraycolsep=0pt
                                                \begin{array}{l}
-                              \text{if $\nonZero{\Phi.\reg(\rv)}$} \\ 
-                              \text{  and $\Phi.reg(lv) = c$} \\
-                              \text{  and $c=(\perm,\start,\addrend,\addr)$}\\
-                              \text{  and $\executeAllowed{\perm}$}
+                                                 \text{if $\nonZero{\memreg(\rv)}$} \\ 
+                                                 \quad\text{and $\memreg(lv) = c$} \\
+                                                 \quad\text{and $c=(\perm,\start,\addrend,\addr)$}\\
+                                                 \quad\text{and $\executeAllowed{\perm}$}
                                                \end{array}
                                                \\ %TODO Maybe combine with jump. (failed + this)
-                            \Phi.\reg\update{\pcreg}{\Phi.\reg(\pcreg) + 1} & \text{if not $\nonZero{\Phi.\reg(\rv)}$}\\
+                                               \updateReg{\pcreg}{\memreg(\pcreg) + 1} & \text{if not $\nonZero{\memreg(\rv)}$}\\
                                                \failed & \text{otherwise }
@@
                                              \begin{cases}
-                                   \Phi.\reg\update{r_1}{\var{w}} &
+                                               \updateReg{r_1}{\var{w}} &
+                                               \arraycolsep=0pt
                                                \begin{array}{l}
-                                     \text{if }\Phi.\reg(r_2) = (\perm,\start,\addrend,\addr)\\
-                                     \text{  and }\readAllowed{\perm} \text{ and } \var{w} = \Phi.\heap(\addr)
+                                                 \text{if } r_1 \neq \pcreg \\
+                                                 \quad\text{and }\memreg(r_2) = (\perm,\start,\addrend,\addr) = \var{c} \\
+                                                 \quad\text{and }\readAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
+                                                 \quad\text{and }\var{w} = \memheap(\addr)
                                                \end{array}\\
@@
                                              \begin{cases}
-                                   \Phi.\heap\update{\addr}{\var{w}} &
+                                               \updateHeap{\addr}{\var{w}} &
+                                               \arraycolsep=0pt
                                                \begin{array}{l}
-                                     \text{if }\Phi.\reg(r_1) = (\perm,\start,\addrend,\addr)\\
-                                     \text{  and }\writeAllowed{\perm} \text{ and } \var{w} = \Phi.\reg(r_2)
+                                                 \text{if }\memreg(r_1) = (\perm,\start,\addrend,\addr) = \var{c} \\
+                                                 \quad\text{and }\writeAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
+                                                 \quad\text{and }\var{w} = \memreg(r_2)
                                                \end{array}\\
@@
                                              \end{cases}\\
- \sem{\move{\refreg{r_1}}{\rv}} & =
+ \sem{\move{\refreg{r_1}}{\rv}}            & = \begin{cases}
+                                                 \updateReg{r_1}{\memreg(\rv)} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.
+                                                 \failed   & \text{otherwise }
+                                 \end{cases}
+\end{align*}
+\begin{align*}
+  \sem{\lea{\refreg{r_1}}{\rv}}            & =
                                              \begin{cases}
-                                   \Phi.\reg\update{r_1}{\Phi.\reg(\rv)} & \text{if $r_1 \neq \pcreg$} \\
-                                   \failed   & \text{otherwise }
+                                               \updateReg{r_1}{\var{c}} &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}{l}
+                                                   \text{if either $n = \rv$ or $\rv = \refheap{r_2}$ and $n = \memreg(r_2)$} \\
+                                                   \quad\text{and in either case $n \in \ints $} \\
+                                                   \quad\text{and $\memreg(r_1) = \stdcap$}\\
+                                                   \quad\text{and $\var{c} = (\perm,\start,\addrend,\addr + n)$}
+                                                 \end{array}\\
+                                               \failed               & \text{otherwise}
+                                             \end{cases} 
+\\ % In the M-Machine, lea checks whether the pointer stay within the allowed range in the pointer.
+  \sem{\restrict{\refreg{r_1}}{\rv_1}{\rv_2}}           & =
+                                             \begin{cases}
+                                               \updateReg{r_1}{\var{c}}  &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}{l}
+                                                   \text{if $r_1 \neq \pcreg$}\\
+                                                   \quad\text{and $\memreg(\rv_1) = c'$}\\
+                                                   \quad\text{and $c' = \stdcap$}\\
+                                                   \quad\text{and either $\rv_2 = n$ or $\memreg(\rv_2) = n$}\\
+                                                   \quad\text{and in either case $n \in \ints$}\\
+                                                   \quad\text{and $\var{newPerm} = \encodePerm(n)$}\\
+                                                   \quad\text{and $\var{newPerm} \sqsubseteq \perm$}\\
+                                                   \quad\text{and $c = (\var{newPerm},\start,\addrend,\addr)$}
+                                                 \end{array}\\
+                                               \failed                   & \text{otherwise}
+                                             \end{cases} \\ 
+  \sem{\plus{\refreg{r_1}}{\rv_1}{\rv_2}}               & =
+                                                          \begin{cases}
+                                                            \updateReg{r_1}{n_1+n_2} &
+                                                            \arraycolsep=0pt
+                                                            \begin{array}{l}
+                                                              \text{if $r_1 \neq \pcreg$}\\
+                                                              \quad\text{and $r_1\neq \pcreg$}\\
+                                                              \quad\text{for $i \in \{1,2\}$}\\
+                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}
+                                                              \qquad\text{and in either case $n_i \in \ints$}
+                                                            \end{array}
                                                           \end{cases}\\
+  \sem{\isptr{\lv}{\rv}} & = \undefined \\ 
+  \sem{\setptr{\lv}{\rv}} & = \undefined \\ 
+  \sem{\subseg{\lv}{\rv}{\rv}} & = \undefined 
 \end{align*}
+
 \section{Examples}
@@
 \]
-Here $\size$ is 8. $i_1 \dots i_7$ refers to the instructions in the following program:
+Here $\size$ is 9. $i_1 \dots i_7$ refers to the instructions in the following program:
 \[
   \begin{array}{r l}
-    i_1 :& \load{\refreg{r_2}}{\size-1} \\
-    i_2 :& \load{\refreg{r_1}}{\refheap{r_2}} \\
-    i_3 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
-    i_4 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
-    i_5 :& \move{\refreg{r_2}}{0} \\
-    i_6 :& \jmp{\refreg{r_0}}
+    i_1 :& \move{\refreg{r_2}}{\refreg{\pcreg}} \\
+    i_2 :& \load{\refreg{r_2}}{\size - 2} \\
+    i_3 :& \load{\refreg{r_1}}{\refheap{r_2}}\\
+    i_4 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
+    i_5 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
+    i_6 :& \move{\refreg{r_2}}{0} \\
+    i_7 :& \jmp{\refreg{r_0}}
   \end{array}
@@
     Addr. & Content \\ \hline
-    1   &  $i_1$ \\ \hline
-    2   &  $i_2$ \\ \hline
-    3   &  $i_3$ \\ \hline
-    4   &  $i_4$ \\ \hline
-    5   &  $i_5$ \\ \hline
-    6   &  $i_6$ \\ \hline
+    0   &  $i_1$ \\ \hline
+    1   &  $i_2$ \\ \hline
+    2   &  $i_3$ \\ \hline
+    3   &  $i_4$ \\ \hline
+    4   &  $i_5$ \\ \hline
+    5   &  $i_6$ \\ \hline
+    5   &  $i_7$ \\ \hline
     7   & Cap. for addr. 8 \\ \hline
@@
 \[
-\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monfun \UPred{\Addrs \parfun \Worlds}))
+\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monfun \UPred{\Addrs \parfun \Words}))
 \]
@@
 \union \\
-                   & \{ \npair{\stdcap[readonly]} \mid \\
+                   & \{ \npair{\stdcap[\readonly]} \mid \\
                    & \quad \exists r \in \RegionName \ldotp \\
@@
                    & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
-                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\Phi\update{a}{v}]{n''}{W''}
+                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
 \union \\
@@
 \end{align*}
-\[
-\stdkr \; : \;  \Worlds \monfun \UPred{\Words}%TODO: What should this be?
-\]
+
+Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
 \begin{align*}
+  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\%TODO: What should this be?
+  \observations (W) \defeq & \{ (\reg,\heap) \mid (\reg,\heap) \diverge \}
+\end{align*}
+\begin{align*}
+  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
   \stdkr(W) \defeq & \{ \npair{c} \mid \\
@@
                      \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \diverge \}
+                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
 \end{align*}
 
-\[
-\stder \; : \;  \Worlds \monfun \UPred{\Words}%TODO: What should this be?
-\]
 \begin{align*}
+  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
   \stder(W) \defeq & \{ \npair{\reg} \mid \\
                    & \quad \forall \npair{c} \in \stdkr(W) \\
-                   & \qquad  \forall \heapSat{\heap}{n}{W} \ldotp \\
-                   & \qquad \quad (\reg\update{r_0}{c},\heap) \diverge
+                   & \qquad  \forall \heapSat[\heap]{n}{W} \ldotp \\
+                   & \qquad \quad (\reg\update{r_0}{c},\heap) \in \observations(W) \}
 \end{align*}
@@
                                 \begin{gathered}
-\exists \heapseg_{rem} \ldotp \Phi.\heap = \heapseg_{rem} \uplus \biguplus_{r\in \dom(W)} \heapseg_r\\
+\exists \heapseg_{rem} \ldotp \memheap = \heapseg_{rem} \uplus \biguplus_{r\in \dom(W)} \heapseg_r\\
 \text{and}\\

### 2016-05-10 capability-machine.tex
Added a standard update of pc to the semantics
Updated ticket machine example to use restrict instruction with correct number of args.
Added the remaining cases missing in the V-relation
Wrote an initial attempt for a lemma about the ticket dispenser.

@@
 
-\newcommand{\stdcap}[1][\perm]{\left(#1,\start,\addrend,\offset \right)}
+\newcommand{\stdcap}[1][\perm]{\left(#1,\start,\addrend,\addr \right)}
 
@@
 \newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}
+\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc(#1)}}
 
@@
                                                 \false & \text{otherwise}
-                                              \end{cases}
+                                              \end{cases} \\
+  \stdUpdatePc{\Phi} &=
+                       \begin{cases}
+                         \updateReg{\pcreg}{\var{newPc}} & 
+                           \arraycolsep=0pt
+                           \begin{array}{l}
+                             \text{if $\memreg(\pcreg) = \stdcap$}\\
+                             \quad\text{and $\var{newPc} = (\perm,\start,\addrend,\addr + 1)$}\\
+                           \end{array} \\
+                           \failed & \text{otherwise}
+                       \end{cases} \\
 \end{align*}
@@
                                                \\ %TODO Maybe combine with jump. (failed + this)
-                                               \updateReg{\pcreg}{\memreg(\pcreg) + 1} & \text{if not $\nonZero{\memreg(\rv)}$}\\
+                                               \stdUpdatePc{\Phi} & \text{if not $\nonZero{\memreg(\rv)}$}\\
                                                \failed & \text{otherwise }
@@
                                              \begin{cases}
-                                               \updateReg{r_1}{\var{w}} &
+                                               \stdUpdatePc{\updateReg{r_1}{\var{w}}} &
                                                \arraycolsep=0pt
@@
                                              \begin{cases}
-                                               \updateHeap{\addr}{\var{w}} &
+                                               \stdUpdatePc{\updateHeap{\addr}{\var{w}}} &
                                                \arraycolsep=0pt
@@
  \sem{\move{\refreg{r_1}}{\rv}}            & = \begin{cases}
-                                                 \updateReg{r_1}{\memreg(\rv)} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.
+                                                 \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.
                                                  \failed   & \text{otherwise }
@@
                                              \begin{cases}
-                                               \updateReg{r_1}{\var{c}} &
+                                               \stdUpdatePc{\updateReg{r_1}{\var{c}}} &
                                                  \arraycolsep=0pt
@@
                                              \begin{cases}
-                                               \updateReg{r_1}{\var{c}}  &
+                                               \stdUpdatePc{\updateReg{r_1}{\var{c}}}  &
                                                  \arraycolsep=0pt
@@
                                                           \begin{cases}
-                                                            \updateReg{r_1}{n_1+n_2} &
+                                                            \stdUpdatePc{\updateReg{r_1}{n_1+n_2}} &
                                                             \arraycolsep=0pt
@@
                                                               \quad\text{for $i \in \{1,2\}$}\\
-                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}
+                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}\\
                                                               \qquad\text{and in either case $n_i \in \ints$}
-                                                            \end{array}
+                                                            \end{array}\\
+                                                            \failed & \text{otherwise}
                                                           \end{cases}\\
@@
 \newcommand{\size}{\var{size}}
+Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$.
 \label{sec:tick-disp}
@@
            & \lea{\refreg{r_1}}{-7} \\
-           & \restrict{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
+           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
            & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
            & \lea{\refreg{r_0}}{4} \\
-           & \restrict{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
+           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
            & \jmp{\codelabel{adv}}
@@
 \end{figure}
+In the JS paper, the ticket dispenser program dereferences the counter and returns the value.
+It does not seem like a similar thing would work here as the adversary can do even more.
+In particular, the adversary can execute a $\halt$ instruction that will cause the machine to halt (successfully).
+The adversary could also choose never to return to the return address, which we set up for him.
+\begin{lemma}
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\exec,\init,\init+26,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then if $i$ was the address returned by the call to $\malloc$, then $\refheap(i+8)$ is even and positive.
+\end{lemma}
+The magic number 26 is the size of the ticet dispenser program (minus one), so the $\Phi$ is assumed to have an execute permission for the area of the heap where the program resides.
 
@@
                    & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
-\union \\
+\union \\ & \dots\text{Continues on next page.}
+\end{align*}
+\begin{align*}
                    & \{ \npair{\stdcap[\exec] = c} \mid \\
@@
                    & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
+                   & \quad \land \\
+                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
+                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
+                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+\union \\
+% For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
+                   & \{ \npair{\stdcap[\entry]} \mid \\
+                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
+                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
+                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\stdcap[\exec]), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+\union \\
+                   & \{ \npair{\stdcap[\exec] = c} \mid \\
+                   & \quad \exists r \in \RegionName \ldotp \\
+                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
+                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
+                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
+                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
+                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
                    & \quad \land \\

### 2016-05-24 capability-machine.tex
Changed the heap satisfaction definition slightly.
Changed the ticket dispenser example to use IO address and updated lemma accordingly

@@
 \newcommand{\parfun}{\rightharpoonup}
+\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\longrightarrow}}
 \newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
@@
 \end{align*}
+It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result.
+As we will see in the semantics, it is assumed that the inital process starts with a readwrite capability for this location.
 \begin{align*}
@@
 \subsection{Ticket Dispenser}
+\label{sec:tick-disp}
 \newcommand{\size}{\var{size}}
+\newcommand{\rio}{r_{io}}
 Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$.
-\label{sec:tick-disp}
+Assume that the register $\rio$ contains a capability for the address in memory where I/O is written to (we assume this address is 0).
 \[
   \begin{array}{r l}
+% Allocate memory for ticket dispenser program
     \init: & \move{\refreg{r_1}}{\size} \\
@@
            & \jmp{\malloc} \\
+% Save ticket dispenser code
            & \store{\refheap{r_1}}{(\encode(i_1))} \\
@@
            & \vdots \\
-           & \store{\refheap{r_1}}{(\encode(i_6))} \\
+           & \store{\refheap{r_1}}{(\encode(i_7))} \\
            & \lea{\refreg{r_1}}{1} \\
+% Save capability for counter
            & \move{\refreg{r_0}}{\refreg{r_1}} \\
@@
            & \lea{\refreg{r_1}}{1} \\
+% Save capability for counter two addresses after jump to adv.
+           & \move{\refreg{r_3}}{\refreg{\pcreg}} \\
+           & \lea{\refreg{r_3}}{17} \\
+           & \store{\refheap{r_3}}{r_0} \\
+           & \move{\refreg{r_3}}{0} \\
+% Save counter
            & \store{\refheap{r_1}}{0} \\
-           & \lea{\refreg{r_1}}{-7} \\
+% Sent pointer to start of ticket dispenser code
+           & \lea{\refreg{r_1}}{-8} \\
+% Set io to -1
+           & \store{\refheap{\rio}}{-1} \\
+% Store io capability just after jump to adv
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
+           & \lea{\refreg{r_0}}{9} \\
+           & \store{\refheap{r_0}}{\refreg{\rio}} \\
+% Overwrite io register
+           & \move{\refreg{\rio}}{0} \\
+% Set up entry pointer to ticketDispenser for adv.
            & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
+% Set up return pointer
            & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
-           & \lea{\refreg{r_0}}{4} \\
+           & \lea{\refreg{r_0}}{5} \\
            & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
-           & \jmp{\codelabel{adv}}
+% jump to adv
+           & \jmp{\codelabel{adv}} \\
+           & 0 \quad (\text{this address will be overwritten})\\
+           & 0 \quad (\text{this address will be overwritten})\\
+% retrieve io capability
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
+           & \lea{r_0}{-2} \\
+           & \load{\refreg{r_1}}{\refheap{r_0}} \\
+% retieve counter capability
+           & \lea{r_0}{1} \\
+           & \load{\refreg{r_2}}{\refheap{r_0}} \\
+% load the value of the counter
+           & \load{\refreg{r_0}}{\refheap{r_2}} \\
+% store the value of the counter to io
+           & \store{\refheap{r_1}}{\refreg{r_0}} \\
+           & \halt
   \end{array}
@@
 The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
@@
 \begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\exec,\init,\init+26,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then if $i$ was the address returned by the call to $\malloc$, then $\refheap(i+8)$ is even and positive.
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\exec,\init,\init+56,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
-The magic number 26 is the size of the ticet dispenser program (minus one), so the $\Phi$ is assumed to have an execute permission for the area of the heap where the program resides.
+The magic number 56 is the size of the above big program (minus one), so the configuration is assumed to have an execute permission for the area of the heap where the program resides.
 
@@
 \[
-\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monfun \UPred{\Addrs \parfun \Words}))
+\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monnefun \UPred{\Addrs \parfun \Words}))
 \]
@@
                    & \quad \forall \npair{c} \in \stdkr(W) \\
-                   & \qquad  \forall \heapSat[\heap]{n}{W} \ldotp \\
-                   & \qquad \quad (\reg\update{r_0}{c},\heap) \in \observations(W) \}
+                   & \qquad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
+                   & \qquad \quad (\reg\update{r_0}{c},\heap \uplus \heap_f) \in \observations(W) \}
 \end{align*}
 
-\begin{definition}[Heap staisfaction/erasure]
+\begin{definition}[Heap satisfaction/erasure]
 \begin{align*}
   \heapSat{n}{W} & & \text{iff} & &
-                                \begin{gathered}
-\exists \heapseg_{rem} \ldotp \memheap = \heapseg_{rem} \uplus \biguplus_{r\in \dom(W)} \heapseg_r\\
-\text{and}\\
-\forall r \in \dom(W) \ldotp \npair{\heapseg_r} \in W(r)(W)
-\end{gathered}
+                                      \arraycolsep=0pt
+                                      \begin{array}{l}
+                                      \exists r: \dom(W) \rightarrow \Heaps \ldotp \\
+                                      \qquad\memheap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
+                                      \quad\text{and}\\
+                                      \qquad\forall \iota \in \dom(W) \ldotp \npair{r(\iota)} \in W(\iota)(W)
+                                      \end{array}\\
 \end{align*}
 \end{definition}
+\begin{lemma}[FTLR]
+  For some configuration 
+\end{lemma}
+
 \section{Related reading}

### 2016-05-30 capability-machine.tex
Changed V relation for execute cap to require to be able to execute from any address in range
added FTLR
Added environment/register relation

@@
 \usepackage{url}
+\usepackage{array}
+\usepackage{arydshln}
+\usepackage{ifthen}
 
 \usepackage{tikz}
+\usepackage{multirow}
 
@@
 \declaretheorem[numbered=yes,name=Definition]{definition}
+
+\newcommand{\forcenewline}{$\phantom{v}$\\}
 
@@
 \newcommand{\malloc}{\codelabel{malloc}}
+\newcommand{\counter}{\codelabel{counter}}
+\newcommand{\iocap}{\codelabel{iocap}}
 
@@
 \newcommand{\contintr}[1]{\intr{K}{#1}}
+\newcommand{\envintr}[1]{\intr{G}{#1}}
 \newcommand{\stdvr}{\valueintr{\asmType}}
 \newcommand{\stder}{\exprintr{\asmType}}
+\newcommand{\stdgr}{\envintr{\asmType}}
 \newcommand{\stdkr}{\contintr{\asmType}}
@@
 \newcommand{\rio}{r_{io}}
+\newcommand{\mrlen}{5cm}
+\newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
+ \left . \vphantom{\begin{array}{l}
+ \ifthenelse{\equal{#1}{1}}{3\\}{
+    \ifthenelse{\equal{#1}{2}}{3\\3\\}{
+    \ifthenelse{\equal{#1}{3}}{3\\3\\3\\}{
+    \ifthenelse{\equal{#1}{4}}{3\\3\\3\\3\\}{
+    \ifthenelse{\equal{#1}{5}}{3\\3\\3\\3\\3\\}{
+    \ifthenelse{\equal{#1}{6}}{3\\3\\3\\3\\3\\3\\}{
+      3\\3\\3\\3\\3\\3\\3\\ %7
+  }}}}}}
+  \end{array}} \right \}}}
+}
+\newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
 Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$.
@@
 \[
-  \begin{array}{r l}
+  \begin{array}{r l l p{\mrlen}}
 % Allocate memory for ticket dispenser program
-    \init: & \move{\refreg{r_1}}{\size} \\
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
-           & \lea{\refreg{r_0}}{3} \\
-           & \jmp{\malloc} \\
+    \init: & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & & \\
+           & \lea{\refreg{r_0}}{3} & & \\
+           & \jmp{\malloc} & & \\
 % Save ticket dispenser code
-           & \store{\refheap{r_1}}{(\encode(i_1))} \\
-           & \lea{\refreg{r_1}}{1} \\
-           & \store{\refheap{r_1}}{(\encode(i_2))} \\
-           & \lea{\refreg{r_1}}{1} \\
-           & \vdots \\
-           & \store{\refheap{r_1}}{(\encode(i_7))} \\
-           & \lea{\refreg{r_1}}{1} \\
+           & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
+           & \lea{\refreg{r_1}}{1} & & \\
+           & \store{\refheap{r_1}}{(\encode(i_2))} & & \\
+           & \lea{\refreg{r_1}}{1} & & \\
+           & \vdots & & \\
+           & \store{\refheap{r_1}}{(\encode(i_7))} & & \\
+           & \lea{\refreg{r_1}}{1} & & \\
 % Save capability for counter
-           & \move{\refreg{r_0}}{\refreg{r_1}} \\
-           & \lea{\refreg{r_0}}{1} \\
-           & \store{\refheap{r_1}}{\refreg{r_0}} \\
-           & \lea{\refreg{r_1}}{1} \\
+           & \move{\refreg{r_0}}{\refreg{r_1}} & \bracket{4} & \annotate{4}{Store a capability for the counter address after the ticket dispenser program} \\
+           & \lea{\refreg{r_0}}{1} & & \\
+           & \store{\refheap{r_1}}{\refreg{r_0}} & & \\
+           & \lea{\refreg{r_1}}{1} & & \\
 % Save capability for counter two addresses after jump to adv.
-           & \move{\refreg{r_3}}{\refreg{\pcreg}} \\
-           & \lea{\refreg{r_3}}{17} \\
-           & \store{\refheap{r_3}}{r_0} \\
-           & \move{\refreg{r_3}}{0} \\
+           & \move{\refreg{r_2}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store a capability for the counter in this code} \\
+           & \lea{\refreg{r_2}}{\counter} & & \\
+           & \store{\refheap{r_2}}{r_0} & & \\
+           & \move{\refreg{r_2}}{0} & \bracket{1} & \annotate{1}{Overwrite capability for counter} \\
 % Save counter
-           & \store{\refheap{r_1}}{0} \\
-% Sent pointer to start of ticket dispenser code
-           & \lea{\refreg{r_1}}{-8} \\
+           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialize counter to 0} \\
+% Change pointer to point to start of ticket dispenser code
+           & \lea{\refreg{r_1}}{-8} & \bracket{1} & \annotate{1}{Capability points to start of td code \footnote{This is also the argument for the adversary}} \\
 % Set io to -1
-           & \store{\refheap{\rio}}{-1} \\
+           & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
 % Store io capability just after jump to adv
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
-           & \lea{\refreg{r_0}}{9} \\
-           & \store{\refheap{r_0}}{\refreg{\rio}} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store the io capability on the stack} \\
+           & \lea{\refreg{r_0}}{\iocap} & & \\
+           & \store{\refheap{r_0}}{\refreg{\rio}} & & \\
 % Overwrite io register
-           & \move{\refreg{\rio}}{0} \\
+           & \move{\refreg{\rio}}{0} & \bracket{1} & \annotate{1}{Overwrite io register} \\
 % Set up entry pointer to ticketDispenser for adv.
-           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} \\
+           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} &\bracket{1} & \annotate{1}{Restrict cap. to td} \\
 % Set up return pointer
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
-           & \lea{\refreg{r_0}}{5} \\
-           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Setup a return pointer for the adversary} \\
+           & \lea{\refreg{r_0}}{5} & & \\
+           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} & & \\
 % jump to adv
-           & \jmp{\codelabel{adv}} \\
-           & 0 \quad (\text{this address will be overwritten})\\
-           & 0 \quad (\text{this address will be overwritten})\\
+           & \jmp{\codelabel{adv}} & \bracket{1} & \annotate{1}{Jump to the adversary} \\
+           & 0 & \bracket{1} & \annotate{1}{Address reserved for io cap.} \\
+           & 0 & \bracket{1} & \annotate{1}{Address reserved for counter cap.} \\
 % retrieve io capability
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} \\
-           & \lea{r_0}{-2} \\
-           & \load{\refreg{r_1}}{\refheap{r_0}} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
+           & \lea{r_0}{-2} & & \\
+           & \load{\refreg{\rio}}{\refheap{r_0}} & & \\
 % retieve counter capability
-           & \lea{r_0}{1} \\
-           & \load{\refreg{r_2}}{\refheap{r_0}} \\
+           & \lea{r_0}{1} & \bracket{2} & \annotate{2}{Retrieve the counter capability} \\
+           & \load{\refreg{r_1}}{\refheap{r_0}} & & \\
 % load the value of the counter
-           & \load{\refreg{r_0}}{\refheap{r_2}} \\
+           & \load{\refreg{r_0}}{\refheap{r_1}} & \bracket{2} & \annotate{2}{Read the counter and write it to io address} \\
 % store the value of the counter to io
-           & \store{\refheap{r_1}}{\refreg{r_0}} \\
+           & \store{\refheap{\rio}}{\refreg{r_0}} & & \\
            & \halt
@@
 \]
-Here $\size$ is 9. $i_1 \dots i_7$ refers to the instructions in the following program:
+Here $\size$ is 9.
+The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities.
+$i_1 \dots i_7$ refers to the instructions in the following program:
 \[
@@
 \begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\exec,\init,\init+56,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+56,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
@@
                    & \quad \land \\
-                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
+                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
+                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}
 \union \\
@@
                    & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\stdcap[\exec]), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
 \union \\
-                   & \{ \npair{\stdcap[\exec] = c} \mid \\
+                   & \{ \npair{\stdcap[\rwx] = c} \mid \\
                    & \quad \exists r \in \RegionName \ldotp \\
@@
                    & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
-                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
+                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\
                    & \quad \land \\
                    & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad \npair[n']{[\pcreg \mapsto c, r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
+                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
 \}
@@
   \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\%TODO: What should this be?
-  \observations (W) \defeq & \{ (\reg,\heap) \mid (\reg,\heap) \diverge \}
+  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid (\reg,\heap) \diverge \}
 \end{align*}
@@
                      \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad ([\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
+                   & \qquad \quad (n,[\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
 \end{align*}
@@
   \stder(W) \defeq & \{ \npair{\reg} \mid \\
-                   & \quad \forall \npair{c} \in \stdkr(W) \\
+                   & \quad \forall \npair{c} \in \stdkr(W) \ldotp \\
                    & \qquad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
-                   & \qquad \quad (\reg\update{r_0}{c},\heap \uplus \heap_f) \in \observations(W) \}
+                   & \qquad \quad (n,\reg\update{r_0}{c},\heap \uplus \heap_f) \in \observations(W) \}
+\end{align*}
+
+\begin{align*}
+  \stdgr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
+  \stdgr(W) \defeq & \{ \npair{\reg} \mid \\
+                    & \quad \forall r \in \RegName \ldotp \\
+                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
 \end{align*}
@@
 \end{align*}
+
 \end{definition}
-\begin{lemma}[FTLR]
-  For some configuration 
+\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
+  If for all $n \in \nats$, $W \in \Worlds$, and register files $\reg$, $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
 \end{lemma}

### 2016-05-31 capability-machine.tex
.

@@
 Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$.
-Assume that the register $\rio$ contains a capability for the address in memory where I/O is written to (we assume this address is 0).
+Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).
 \[
@@
 \begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+56,\init)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
-The magic number 56 is the size of the above big program (minus one), so the configuration is assumed to have an execute permission for the area of the heap where the program resides.
+The variable $\var{proglen}$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
 
@@
 \end{align*}
+``Continuation'' relation:
 \begin{align*}
@@
 \end{align*}
-
+``Expression'' relation:
 \begin{align*}
@@
 \end{align*}
-
+Register-file relation:
 \begin{align*}

### 2016-05-31 capability-machine.tex
attempt at new observation relation
Added a small description of the observation relation

@@
 \newcommand{\diverge}[1][n]{\not\Downarrow_{#1}}
+\newcommand{\step}[1][]{\rightarrow_{#1}}
 
@@
 \begin{align*}
-  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\%TODO: What should this be?
-  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid (\reg,\heap) \diverge \}
+  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
+  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
+                           & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
+                           & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
+                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
+                           & \qquad \Rightarrow \npair[n-1]{\heap(0)} \in \stdvr(W) \\
+                           & \quad \qquad \land \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap(0)} \in W(r)(W) ))\}
 \end{align*}
+The ``good observations'' are as follows
+\begin{itemize}
+\item The execution does not stop in $n$ steps. 
+\item The execution stops with a $\failed$.
+\item The execution halts in at most $n$ steps with a valid value in in address 0 and the value at address 0 is also valid with respect to the world.
+\end{itemize}
+Address 0 is the designated I/O address.
+
 ``Continuation'' relation:

### 2016-05-31 capability-machine.tex
.
typo

@@
 \begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,1,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
@@
   \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
+00
+
                      \union \\
@@
                            & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
-                           & \qquad \Rightarrow \npair[n-1]{\heap(0)} \in \stdvr(W) \\
-                           & \quad \qquad \land \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap(0)} \in W(r)(W) ))\}
+                           & \qquad \Rightarrow \npair[n-1]{\heap(0)} \in \stdvr(W) \land \\
+                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap(0)} \in W(r)(W) )\}
 \end{align*}

### 2016-05-31 capability-machine.tex
typos

@@
   \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
-00
-
 \union \\
@@
                            & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
-                           & \qquad \Rightarrow \npair[n-1]{\heap(0)} \in \stdvr(W) \land \\
+                           & \qquad \Rightarrow \npair[n-i]{\heap(0)} \in \stdvr(W) \land \\
                            & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap(0)} \in W(r)(W) )\}
@@
   \stdkr(W) \defeq & \{ \npair{c} \mid \\
-                   & \quad \forall v\ldotp \in \stdvr(W) \Rightarrow \\
+                   & \quad \forall v \in \stdvr(W)\ldotp \\
                    & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.

### 2016-06-08 capability-machine.tex
Changed some of the definition for the value relation.
Updated the value interpretation definition and factored out common parts in the definition
typo

@@
 %Math notation
+\newcommand{\restrictfun}[1]{|_{#1}}
 \newcommand{\parfun}{\rightharpoonup}
@@
 \newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
+\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
+\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
+\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
 \newcommand{\union}{\mathbin{\cup}}
@@
 \newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
+
 \newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
@@
 \newcommand{\stdUpdatePc}[1]{\plainfun{updatePc(#1)}}
+
+\newcommand{\readCond}[1]{\plainfun{readCondition}(#1)}
+\newcommand{\writeCond}[1]{\plainfun{writeCondition}(#1)}
+\newcommand{\execCond}[1]{\plainfun{executeCondition}(#1)}
+\newcommand{\entryCond}[1]{\plainfun{entryCondition}(#1)}
 
@@
 \[
+  \iota_{\var{start},\var{end}} \; : \; \Worlds \monfun \UPred{\Addrs \parfun \Heaps}
+\]
+\begin{align*}
+  \iota_{\var{start},\var{end}}(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
+                                     & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
+\end{align*}
+
+\begin{align*}
+  \readCond{n,W,\start,\addrend} = & \;\exists r \in \RegionName \ldotp \\
+                                   & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
+  \writeCond{n,W,\start,\addrend} = & \; \exists r \in \RegionName \ldotp \\
+                                   & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\
+  \execCond{n,W,\start,\addrend} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
+                                   & \; \quad \forall \npair[n']{v_1},\dots,\npair[n']{v_k} \in \stdvr(W'),\addr' \in [\start, \addrend] \ldotp \\
+                                   & \; \qquad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \\
+                                   & \; \qquad \quad \in \stder(W') \\
+  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
+                                   & \; \quad \forall \npair[n']{v_1},\dots,\npair[n']{v_k} \in \stdvr(W') \ldotp \\
+                                   & \; \qquad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \\
+                                   & \; \qquad \quad \in \stder(W') \\
+\end{align*}
+
+\[
 \stdvr \; : \;  \Worlds \monfun \UPred{\Words}
@@
 \union \\
-                   & \{ \npair{\stdcap[\readonly]} \mid \\
-                   & \quad \exists r \in \RegionName \ldotp \\
-                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
-\union \\
-                   & \{ \npair{\readwrite} \mid \\
-                   & \quad \exists r \in \RegionName \ldotp \\
-                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
-                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
-                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
-                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
-                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
-\union \\ & \dots\text{Continues on next page.}
-\end{align*}
-\begin{align*}
-                   & \{ \npair{\stdcap[\exec] = c} \mid \\
-                   & \quad (\exists r \in \RegionName \ldotp \\
-                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
-                   & \quad \land \\
-                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
-                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
-                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}
+                   & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
+%                   & \quad \exists r \in \RegionName \ldotp \\
+%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
+                   & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
+%                   & \quad \exists r \in \RegionName \ldotp \\
+%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
+%                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
+%                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
+%                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
+%                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
+                   & \{ \npair{\stdcap[\exec]} \mid \\
+                   & \quad\readCond{n,W,\start,\addrend} \land \\
+%                   & \quad (\exists r \in \RegionName \ldotp \\
+%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
+                   & \quad \execCond{n,W,\start,\addrend} \}
+%                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
+%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
+%                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}
 \union \\
 % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
-                   & \{ \npair{\stdcap[\entry]} \mid \\
-                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
-\union \\
+                   & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
+%                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
+%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
+%                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
                    & \{ \npair{\stdcap[\rwx] = c} \mid \\
-                   & \quad \exists r \in \RegionName \ldotp \\
-                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
-                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
-                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
-                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
-                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\
-                   & \quad \land \\
-                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
-                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
-\}
+                   & \quad \writeCond{n,W,\start,\addrend} \land \\
+                   & \quad \execCond{n,W,\start,\addrend} \}
+%                   & \quad \exists r \in \RegionName \ldotp \\
+%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
+%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
+%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
+%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
+%                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
+%                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
+%                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
+%                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\
+%                   & \quad \land \\
+%                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
+%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
+%                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
 \end{align*}
@@
                            & \qquad \Rightarrow \npair[n-i]{\heap(0)} \in \stdvr(W) \land \\
-                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap(0)} \in W(r)(W) )\}
+                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap \restrictfun{\{0\}}} \in W(r)(W) )\}
 \end{align*}
@@
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  If for all $n \in \nats$, $W \in \Worlds$, and register files $\reg$, $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
+  For any $n \in \nats$, $W \in \Worlds$, and register file $\reg$, if  $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
 \end{lemma}

### 2016-06-08 capability-machine.tex
Make the document compile with my pdflatex.
Updated reg-file relation and expression relation to only consider pc. Updated the value relation according
Updated the FTLR to be consistent with update to expression relation

@@
 \usepackage{ifthen}
+\usepackage{ifpdf}
 
@@
   \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionName \ldotp \\
-                                   & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\
-  \execCond{n,W,\start,\addrend} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                   & \; \quad \forall \npair[n']{v_1},\dots,\npair[n']{v_k} \in \stdvr(W'),\addr' \in [\start, \addrend] \ldotp \\
-                                   & \; \qquad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \\
-                                   & \; \qquad \quad \in \stder(W') \\
+                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
+  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
+                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
+                                          & \; \qquad \npair{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
   \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                   & \; \quad \forall \npair[n']{v_1},\dots,\npair[n']{v_k} \in \stdvr(W') \ldotp \\
-                                   & \; \qquad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \\
-                                   & \; \qquad \quad \in \stder(W') \\
+                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
 \end{align*}
@@
                    & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
-%                   & \quad \exists r \in \RegionName \ldotp \\
-%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)\} 
                    & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
-%                   & \quad \exists r \in \RegionName \ldotp \\
-%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
-%                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
-%                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
-%                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
-%                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \}
                    & \{ \npair{\stdcap[\exec]} \mid \\
                    & \quad\readCond{n,W,\start,\addrend} \land \\
-%                   & \quad (\exists r \in \RegionName \ldotp \\
-%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W)) \\
-                   & \quad \execCond{n,W,\start,\addrend} \}
-%                   & \quad (\forall n' < n \ldotp \forall W' \future W \ldotp \\
-%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
-%                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W')) \}
+                   & \quad \execCond{n,W,\start,\addrend,\exec} \}
 \union \\
@@
                    & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
-%                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
-%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W') \ldotp \\
-%                   & \qquad \quad \npair[n']{[\pcreg \mapsto \stdcap[\exec], r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
                    & \{ \npair{\stdcap[\rwx] = c} \mid \\
                    & \quad \writeCond{n,W,\start,\addrend} \land \\
-                   & \quad \execCond{n,W,\start,\addrend} \}
-%                   & \quad \exists r \in \RegionName \ldotp \\
-%                   & \qquad  \forall \addr \in \{\start, \dots, \addrend \} \ldotp \\
-%                   & \qquad \quad  \forall \npair{\heapseg} \in W(r)(W)\ldotp \\
-%                   & \qquad \qquad   \addr \in \dom(\heapseg) \land \\
-%                   & \qquad \qquad   \npair[n-1]{\heapseg(\addr)} \in \stdvr(W) \land \\
-%                   & \qquad \qquad   \forall n' < n, W' \future W, v \ldotp \\
-%                   & \qquad \qquad \quad \heapSat{n'}{W'} \land \\
-%                   & \qquad \qquad \quad \npair[n']{v} \in \stdvr(W') \\
-%                   & \qquad \qquad \qquad  \Rightarrow \exists W'' \future W' \ldotp \heapSat[\memheap\update{a}{v}]{n''}{W''} \\
-%                   & \quad \land \\
-%                   & \quad \forall n' < n \ldotp \forall W' \future W \ldotp \\
-%                   & \qquad \forall \npair[n']{v_1,\dots,v_k} \in \stdvr(W'),\addr' \in \{\start, \dots, \addrend \} \ldotp \\
-%                   & \qquad \quad \npair[n']{[\pcreg \mapsto (\exec,\start,\addrend,\addr'), r_1 \mapsto v_1, \dots , r_k \mapsto v_k]} \in \stder(W') \}
+                   & \quad \execCond{n,W,\start,\addrend,\exec} \land \\
+                   & \quad \execCond{n,W,\start,\addrend,\rwx} \}
 \end{align*}
@@
 
+Register-file relation:
+\begin{align*}
+  \stdgr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
+  \stdgr(W) \defeq & \{ \npair{\reg} \mid \\
+                    & \quad \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
+\end{align*}
+
 ``Continuation'' relation:
@@
                    & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat[\heap]{n'}{W'}, \npair{v_1 \dots v_k} \in \stdvr(W') \ldotp \\
-                   & \qquad \quad (n,[\pcreg \mapsto \updatePcPerm{c},r_0 \mapsto v,r_1 \mapsto v_1,\dots, r_k \mapsto v_k], \heap) \in \observations(W) \}
+                     \forall \heapSat[\heap]{n'}{W'} \ldotp \forall \reg \in \stdgr(W') \\
+                   & \qquad \quad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
 \end{align*}
@@
 \begin{align*}
-  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
-  \stder(W) \defeq & \{ \npair{\reg} \mid \\
-                   & \quad \forall \npair{c} \in \stdkr(W) \ldotp \\
-                   & \qquad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
-                   & \qquad \quad (n,\reg\update{r_0}{c},\heap \uplus \heap_f) \in \observations(W) \}
-\end{align*}
-Register-file relation:
-\begin{align*}
-  \stdgr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
-  \stdgr(W) \defeq & \{ \npair{\reg} \mid \\
-                    & \quad \forall r \in \RegName \ldotp \\
-                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
+  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
+  \stder(W) \defeq & \{ \npair{\pc} \mid \\
+                   & \quad \forall \reg \in \stdgr(W) \ldotp \\
+                   & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\
+                   & \qquad \quad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
+                   & \qquad \qquad (n,\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f) \in \observations(W) \}
 \end{align*}
@@
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  For any $n \in \nats$, $W \in \Worlds$, and register file $\reg$, if  $\npair{\reg} \in \stdgr(W)$, then $(n,\reg) \in \stder(W)$
+  For any $n \in \nats$, $W \in \Worlds$, and a word $\pc$, if  $\npair{\pc} \in \stdvr(W)$, then $\npair{\pc} \in \stder(W)$
 \end{lemma}

### 2016-06-08 capability-machine.tex
small mistake in tck disp lemma

@@
 \begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,1,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}

### 2016-06-13 capability-machine.tex
small changes and typos

@@
 \newcommand{\Worlds}{\plaindom{World}}
+\newcommand{\StorePred}{\plaindom{StorePred}}
 \newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
@@
 \begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
@@
 \[
-\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monnefun \UPred{\Addrs \parfun \Words}))
+  \StorePred = \Worlds \monnefun \UPred{\Addrs \parfun \Words}
+\]
+\[
+\Worlds \cong \frac{1}{2} (\RegionName \parfun \StorePred)
 \]
@@
 \[
-  \iota_{\var{start},\var{end}} \; : \; \Worlds \monfun \UPred{\Addrs \parfun \Heaps}
+  \iota_{\var{start},\var{end}} \; : \; \StorePred
 \]
@@
                                           & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
-                                          & \; \qquad \npair{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
+                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
   \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
@@
                    & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
-                   & \{ \npair{\stdcap[\rwx] = c} \mid \\
+                   & \{ \npair{\stdcap[\rwx]} \mid \\
                    & \quad \writeCond{n,W,\start,\addrend} \land \\
@@
                            & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
-                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
-                           & \qquad \Rightarrow \npair[n-i]{\heap(0)} \in \stdvr(W) \land \\
-                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap \restrictfun{\{0\}}} \in W(r)(W) )\}
+                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap')  \\
+                           & \qquad \Rightarrow \npair[n-i]{\heap'(0)} \in \stdvr(W) \land \\
+                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap' \restrictfun{\{0\}}} \in W(r)(W) )\}
 \end{align*}

### 2016-06-14 capability-machine.tex
Added simple intrastructure for comments. Added specification of start of proof.

@@
 \declaretheorem[numbered=yes,name=Definition]{definition}
+\declaretheorem[numbered=yes,name=Specification]{specification}
 
@@
 \newcommand{\true}{\mathit{true}}
+
+%Comments
+\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}}
+\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}}
 
@@
 \newcommand{\contintr}[1]{\intr{K}{#1}}
-\newcommand{\envintr}[1]{\intr{G}{#1}}
+\newcommand{\regintr}[1]{\intr{R}{#1}}
 \newcommand{\stdvr}{\valueintr{\asmType}}
 \newcommand{\stder}{\exprintr{\asmType}}
-\newcommand{\stdgr}{\envintr{\asmType}}
+\newcommand{\stdrr}{\regintr{\asmType}}
 \newcommand{\stdkr}{\contintr{\asmType}}
@@
 Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).
+The following is a test program for a ticket dispenser:
 \[
   \begin{array}{r l l p{\mrlen}}
+% Set io to -1
+\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
+% Store io capability just after jump to adv
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store the io capability on the stack} \\
+           & \lea{\refreg{r_0}}{\iocap} & & \\
+           & \store{\refheap{r_0}}{\refreg{\rio}} & & \\
+% Overwrite io register
+           & \move{\refreg{\rio}}{0} & \bracket{1} & \annotate{1}{Overwrite io register} \\
 % Allocate memory for ticket dispenser program
-    \init: & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\
+     & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\
            & \move{\refreg{r_0}}{\refreg{\pcreg}} & & \\
@@
            & \lea{\refreg{r_1}}{-8} & \bracket{1} & \annotate{1}{Capability points to start of td code \footnote{This is also the argument for the adversary}} \\
-% Set io to -1
-           & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
-% Store io capability just after jump to adv
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store the io capability on the stack} \\
-           & \lea{\refreg{r_0}}{\iocap} & & \\
-           & \store{\refheap{r_0}}{\refreg{\rio}} & & \\
-% Overwrite io register
-           & \move{\refreg{\rio}}{0} & \bracket{1} & \annotate{1}{Overwrite io register} \\
 % Set up entry pointer to ticketDispenser for adv.
@@
 The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities.
-$i_1 \dots i_7$ refers to the instructions in the following program:
+$i_1 \dots i_7$ refers to the instructions in the following ticket dispenser:
 \[
@@
 The adversary could also choose never to return to the return address, which we set up for him.
-\begin{lemma}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+\begin{lemma}[Ticket dispenser test]
+\label{lem:tckt-disp}
+ Given $\Phi$, if $\memheap$ contains the ticket dispenser test program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
 The variable $\var{proglen}$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
+
+The ticket dispenser test program uses $\malloc$.
+To be able to reason about programs that use $\malloc$, we assume that it obeys the following specification.
+If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
+\begin{specification}[Malloc]
+  For all worlds, $W$, configurations, $\Phi$, natural numbers, $n$, and addresses $b$, $e$, and $return$ where $\heapSat[\memheap]{n}{W}$ if
+  \begin{itemize}
+  \item $\memreg(r_0) = (\entry,b,e,\var{return})$,
+  \item $\memreg(r_1) = m$ where $m$ is some positive natural number\footnote{We do not include 0 as malloc in C returns a null pointer when malloc(0) is called.}, and
+  \item the remaining registers contain 0,
+  \end{itemize}
+then there exists a configuration $\Phi'$, a future world $W' \future W$, a natural number $i$ such that $\heapSat[{\memheap[\Phi']}]{n-i}{W'}$ and $\dom(\memheap[\Phi']) \supseteq \dom(\memheap) \uplus [b',e']$\footnote{notice that $\malloc$ is allowed to allocated memory for its own purposes} for some addresses $b'$ and $e'$.
+Further the register file contains the following
+\begin{itemize}
+\item $\memreg(\pcreg) = (\rwx,b,e,\var{return})$,
+\item $\memreg(r_0) = (\entry,b,e,\var{return})$,
+\item $\memreg(r_1) = (\rwx,b',e',b')$, and
+\item $\memreg(r) = 0$ for all other registers $r$.
+\end{itemize}
+\end{specification}
+\begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
+Let configuration $\Phi$ be given. Assume that the ticket dispenser test code resides in memory starting at address $\init$. Further assume that the register file contains the following:
+\begin{enumerate}
+\item $\memreg(\rio) = (\readwrite,0,0,0)$ 
+\item $\memreg(\pcreg) = (\rwx,\init, \init + \var{proglen} - 1, \init)$
+\item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
+\end{enumerate}
+Further assume $\Phi \rightarrow^* (\halted,\Phi')$.
+\end{proof}
 
@@
 \begin{align*}
-  \stdgr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
-  \stdgr(W) \defeq & \{ \npair{\reg} \mid \\
+  \stdrr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
+  \stdrr(W) \defeq & \{ \npair{\reg} \mid \\
                     & \quad \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
@@
                    & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat[\heap]{n'}{W'} \ldotp \forall \reg \in \stdgr(W') \\
+                     \forall \heapSat[\heap]{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
                    & \qquad \quad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
@@
   \stder(W) \defeq & \{ \npair{\pc} \mid \\
-                   & \quad \forall \reg \in \stdgr(W) \ldotp \\
+                   & \quad \forall \reg \in \stdrr(W) \ldotp \\
                    & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\

### 2016-06-15 capability-machine.tex
work in progress

@@
 \newcommand{\var}[1]{\mathit{#1}}
+\newcommand{\hs}{\var{hs}}
 \newcommand{\hv}{\var{hv}}
@@
 \newcommand{\Heaps}{\plaindom{Heap}}
+\newcommand{\HeapSegments}{\plaindom{HeapSegment}}
 \newcommand{\Confs}{\plaindom{Conf}}
@@
 \Caps  &::= \Perms \times \Addrs \times \Addrs \times \Addrs & & &
-\Confs &::= \Mems + \{\failed, \halted \times \Mems\}
+\Confs &::= \Mems + \{\failed, \halted \times \Mems\} \\
+\HeapSegments &::= \Addrs \parfun \Words & & & &
 \end{align*}
@@
 Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).
+We assume that entry capabilities for $\codelabel{adv}$ and $\malloc$ is ambiently available.
 The following is a test program for a ticket dispenser:
@@
 \label{lem:tckt-disp}
- Given $\Phi$, if $\memheap$ contains the ticket dispenser test program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
+ Given a configuration $\Phi$, if
+ \begin{description}
+ \item[$\boldsymbol \memheap$] contains the ticket dispenser test program starting at the address labelled $\init$, 
+ \item[$\boldsymbol \memheap$] contains the adversary code starting at label $\codelabel{adv}$.
+The adversary is assumed to have no capabilities on the heap, but it has access to the ambient $\malloc$ capability!
+\lau{maybe mention that malloc is also on the heap}
+ \item[the heap segment $\boldsymbol{ \hs}$] contains all allocated addresses in the sense that any capability in $\Phi$ is a capability for a piece of memory in $\hs$,
+ \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\init,\init+\var{proglen}-1,\init)$, which is the only capability that grants any access to this part of the heap,
+ \item[$\boldsymbol{ \memreg(\rio)}$] is $(\readwrite,0,0,0)$, which also is the only capability that grants access to this part of the heap, 
+ \item[the remaining registers] all contains 0,
+ \end{description}
+ and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
 \end{lemma}
@@
   \end{itemize}
-then there exists a configuration $\Phi'$, a future world $W' \future W$, a natural number $i$ such that $\heapSat[{\memheap[\Phi']}]{n-i}{W'}$ and $\dom(\memheap[\Phi']) \supseteq \dom(\memheap) \uplus [b',e']$\footnote{notice that $\malloc$ is allowed to allocated memory for its own purposes} for some addresses $b'$ and $e'$.
+then there exists a configuration $\Phi'$, a future world $W[r \mapsto \iota_{b',e'}] \future W$ for some addresses $b'$ and $e'$\lau{I had this to be exists future world $W'$, but it seemed useless as I couldn't argue that I could change the heap}, a natural number $i$ such that $\heapSat[{\memheap[\Phi']}]{n-i}{W'}$, $\dom(\memheap[\Phi']) = \dom(\memheap) \uplus [b',e']$  and for all $a \in \dom(\memheap)$ the heap is unchanged, i.e., $\memheap(a) = \memheap[\Phi'](a)$\footnote{This is probably not the best assumption, as $\malloc$ probably need to keep track of the allocated memory, but this is a simple way to make $\malloc$ trustworthy in the sense that it cannot give the same capability to the adversary code.}.
 Further the register file contains the following
@@
 \begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
-Let configuration $\Phi$ be given. Assume that the ticket dispenser test code resides in memory starting at address $\init$. Further assume that the register file contains the following:
+Let configuration $\Phi$ be given.
+Assume that the ticket dispenser test code resides in memory starting at address $\init$ along with the adversary code at address $\codelabel{adv}$.
+Also assume that capabilities for $\codelabel{adv}$ and $\malloc$ are ambiently available.
+Further assume that the register file contains the following:
 \begin{enumerate}
@@
 Further assume $\Phi \rightarrow^* (\halted,\Phi')$.
+As the evaluation terminates, this must happen in a finite number of steps, say this number is $n$.
+We also assume that there is a heap segment $\hs$ as described in the lemma.
+If we execute $\Phi$ until just before the jump to $\malloc$, then a number of steps has occured, say $0<i<n$, further location 0 on the heap has been overwritten with $-1$, the io capability has been saved to the designated address on the heap, and the $\rio$ register has been overwritten with 0.
+
+Now we need to jump to $\malloc$, so we need to make sure that the precondition in the specification is satisfied.
+To do this, we first need a world that models the part of the heap that is currently in use which is $\hs$.
+We pick the world $W$ to be:
+\begin{align*}
+  W(\codelabel{adv}) & = \iota_{\var{adv},\var{advend}} \\
+  W(\init)           & = \iota_{\init,\init+\var{proglen} - 1} \\
+  W(\rio)            & = \lambda W\ldotp \{(n,\hs) \mid \hs(0) = -1 \text{ or } \\
+                     &   \hphantom{ = \lambda W\ldotp \{(n,\hs) \mid }\:\; \hs(0) \text{ even and positive} \}
+\end{align*}
+for all other addresses $a$ that are not captured in the above but are part of $hs$, we define a new region $r$ in $W$ where
+\begin{align*}
+  W(r) & = \iota_{a,a}
+\end{align*}
+Then we need to argue that $\heapSat{\hs}{n-i}{W}$.
+Given any address, in $\hs$ there is a region that covers that address, and the regions cover only the addresses in $\hs$.
+Now each of the heap segments have to be accepted by the respective regions.
+For the adversary, we assumed that it had no capabilities, so all the addresses contain intergers which i trivially in the value relation.
+The same goes for the heap segment with the ticket dispenser test program and the remaining addresses of the heap but addresses 0 \lau{need assumption about this.} The io address contains $-1$ at this point of the execution which mean that the heapsegment only consisting of address 0 is accepted in this world.
+
+The remaining conditions in the specification are also satisfied, so there exists a configuration $\Phi'$
 \end{proof}
@@
                    & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat[\heap]{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
+                     \forall \heapSat{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
                    & \qquad \quad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
@@
                    & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\
-                   & \qquad \quad  \forall \heapSat[\heap]{n}{W}, \heap_f \ldotp \\
+                   & \qquad \quad  \forall \heapSat{n}{W}, \heap_f \ldotp \\
                    & \qquad \qquad (n,\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f) \in \observations(W) \}
@@
                                       \begin{array}{l}
-                                      \exists r: \dom(W) \rightarrow \Heaps \ldotp \\
-                                      \qquad\memheap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
+                                      \exists r: \dom(W) \rightarrow \HeapSegments \ldotp \\
+                                      \qquad\heap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
                                       \quad\text{and}\\

### 2016-06-15 capability-machine.tex
changed the malloc specification

@@
 \begin{specification}[Malloc]
-  For all worlds, $W$, configurations, $\Phi$, natural numbers, $n$, and addresses $b$, $e$, and $return$ where $\heapSat[\memheap]{n}{W}$ if
+  For all worlds, $W$, heap segments, $\hs$, register files ,$\reg$, natural numbers, $n>j$ ($j$ is the number of steps the allocation runs for), and addresses $b$, $e$, and $return$ where $\heapSat[\hs]{n}{W}$ if
   \begin{itemize}
   \item $\memreg(r_0) = (\entry,b,e,\var{return})$,
-  \item $\memreg(r_1) = m$ where $m$ is some positive natural number\footnote{We do not include 0 as malloc in C returns a null pointer when malloc(0) is called.}, and
+  \item $\memreg(r_1) = m$ where $m>0$\footnote{We do not include 0 as malloc in C returns a null pointer when malloc(0) is called.}, and
   \item the remaining registers contain 0,
   \end{itemize}
-then there exists a configuration $\Phi'$, a future world $W[r \mapsto \iota_{b',e'}] \future W$ for some addresses $b'$ and $e'$\lau{I had this to be exists future world $W'$, but it seemed useless as I couldn't argue that I could change the heap}, a natural number $i$ such that $\heapSat[{\memheap[\Phi']}]{n-i}{W'}$, $\dom(\memheap[\Phi']) = \dom(\memheap) \uplus [b',e']$  and for all $a \in \dom(\memheap)$ the heap is unchanged, i.e., $\memheap(a) = \memheap[\Phi'](a)$\footnote{This is probably not the best assumption, as $\malloc$ probably need to keep track of the allocated memory, but this is a simple way to make $\malloc$ trustworthy in the sense that it cannot give the same capability to the adversary code.}.
-Further the register file contains the following
+then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat{\hs'}{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ for some register file $\reg'$ that has the following constraints:
 \begin{itemize}
-\item $\memreg(\pcreg) = (\rwx,b,e,\var{return})$,
-\item $\memreg(r_0) = (\entry,b,e,\var{return})$,
-\item $\memreg(r_1) = (\rwx,b',e',b')$, and
-\item $\memreg(r) = 0$ for all other registers $r$.
+\item $\reg'(\pcreg) = (\rwx,b,e,\var{return})$,
+\item $\reg'(r_0) = (\entry,b,e,\var{return})$,
+\item $\reg'(r_1) = (\rwx,b',e',b')$, and
+\item $\reg'(r) = 0$ for all other registers $r$.
 \end{itemize}

### 2016-06-15 capability-machine.tex
typo

@@
   \end{itemize}
-then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat{\hs'}{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ for some register file $\reg'$ that has the following constraints:
+then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat[\hs']{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ for some register file $\reg'$ that has the following constraints:
 \begin{itemize}

### 2016-06-15 capability-machine.tex
added a new specification for malloc based on Dominiques advise.

@@
 If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
-\begin{specification}[Malloc]
+\begin{specification}[Malloc (D)]
+Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.
+If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$.
+Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$ for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$.
+\lau{at what step should this hold?
+Also it should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}
+\end{specification}
+\begin{specification}[Malloc (L)]
   For all worlds, $W$, heap segments, $\hs$, register files ,$\reg$, natural numbers, $n>j$ ($j$ is the number of steps the allocation runs for), and addresses $b$, $e$, and $return$ where $\heapSat[\hs]{n}{W}$ if
@@
   \end{itemize}
-then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat[\hs']{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ for some register file $\reg'$ that has the following constraints:
+then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat[\hs']{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ \lau{need to extend $\hs$ to entire heap} for some register file $\reg'$ that has the following constraints:
 \begin{itemize}
@@
 \end{specification}
+
 \begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
@@
   \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
+                                          & \; \quad \npair[n']{(\rwx,\start,\addrend,\addr)} \in \stder(W') \\
 \end{align*}

### 2016-06-16 capability-machine.tex
removed old malloc specification

@@
 If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$.
-Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$ for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$.
+Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$ \lau{doesn't this invariant allow the newly allocated memory to contain capabilities?} for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$.
 \lau{at what step should this hold?
+Just some step n'<n?
 Also it should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}
-\end{specification}
-\begin{specification}[Malloc (L)]
-  For all worlds, $W$, heap segments, $\hs$, register files ,$\reg$, natural numbers, $n>j$ ($j$ is the number of steps the allocation runs for), and addresses $b$, $e$, and $return$ where $\heapSat[\hs]{n}{W}$ if
-  \begin{itemize}
-  \item $\memreg(r_0) = (\entry,b,e,\var{return})$,
-  \item $\memreg(r_1) = m$ where $m>0$\footnote{We do not include 0 as malloc in C returns a null pointer when malloc(0) is called.}, and
-  \item the remaining registers contain 0,
-  \end{itemize}
-then there exists addresses $b'$ and $e'$ where $b'-e' = m$ such that for all future worlds $W' \future W$ where $\heapSat[\hs']{n-j}{W'}$ for $\hs' = hs[b' \mapsto 0, \dots, e' \mapsto 0]$ we have $(\reg,\hs) \step[j] (\reg',\hs')$ \lau{need to extend $\hs$ to entire heap} for some register file $\reg'$ that has the following constraints:
-\begin{itemize}
-\item $\reg'(\pcreg) = (\rwx,b,e,\var{return})$,
-\item $\reg'(r_0) = (\entry,b,e,\var{return})$,
-\item $\reg'(r_1) = (\rwx,b',e',b')$, and
-\item $\reg'(r) = 0$ for all other registers $r$.
-\end{itemize}
 \end{specification}

### 2016-06-16 capability-machine.tex
added questions for malloc specification

@@
 Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.
-If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$.
-Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$ \lau{doesn't this invariant allow the newly allocated memory to contain capabilities?} for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$.
-\lau{at what step should this hold?
-Just some step n'<n?
-Also it should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}
+If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ \\
+\lau{Is $hs_{\var{frame}}$ supposed to be the rest of memory, so $h$ is something we can execute in?
+Is $\malloc$ allowed to allocate more memory, i.e., how does $\hs_{\var{footprint}}$ relate to $\hs_{\var{footprint}}'$?}\\
+that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$.
+Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$\\
+\lau{doesn't this invariant allow the newly allocated memory to contain capabilities?}\\
+ for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$. 
+\lau{at what step should this hold? Just some step $n'<n$?}\\
+\lau{It should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}\\
+\lau{How do we express that $\malloc$ can be ``trusted'' in the sense that it won't save the capab}
 \end{specification}

### 2016-06-16 capability-machine.tex
Added question to lemma

@@
 \end{lemma}
+\lau{in the above lemma, what assumptions should we make on the adversary?
+In the js paper, we assume syntactically well-formed, do we here assume that the capability for the adversary is in the value relation?
+If so in what world should this be true?
+The one with just $\iota_{\var{adv},\var{adv end}}$?}
 The variable $\var{proglen}$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
@@
 If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
-\begin{specification}[Malloc (D)]
+\begin{specification}[Malloc]
 Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.

### 2016-06-16 capability-machine.tex
added comment question about continuation predicate
added another comment question about continuation predicate

@@
 \end{align*}
-Then we need to argue that $\heapSat{\hs}{n-i}{W}$.
+Then we need to argue that $\heapSat[\hs]{n-i}{W}$.
 Given any address, in $\hs$ there is a region that covers that address, and the regions cover only the addresses in $\hs$.
@@
 \end{align*}
+\lau{$r_1$ contains the ``return value'', shouldn't that be the one we quantify over?} \\
+\lau{When continuation is invoked, it has to be in registers, but here we do not require it to be in the value relation, so it seems impossible to show some of the required conditions.}
 ``Expression'' relation:

### 2016-06-17 capability-machine.tex
update to specification for malloc, made return and invoke more precise

--- 2016-06-16 capability-machine.tex
@@
 It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result.
-As we will see in the semantics, it is assumed that the inital process starts with a readwrite capability for this location.
+As we will see in the semantics, it is assumed that the initial process starts with a read-write capability for this location.
 \begin{align*}
@@
 % Set io to -1
-\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
+\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialise io to $-1$} \\
 % Store io capability just after jump to adv
@@
 % Save counter
-           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialize counter to 0} \\
+           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialise counter to 0} \\
 % Change pointer to point to start of ticket dispenser code
@@
 % retrieve io capability
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capability} \\
            & \lea{r_0}{-2} & & \\
@@
 Here $\size$ is 9.
-The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities.
+The variables $\counter$ and $\iocap$ are respectively the offsets to the addresses reserved for the counter and io capabilities.
 $i_1 \dots i_7$ refers to the instructions in the following ticket dispenser:
@@
 \lau{in the above lemma, what assumptions should we make on the adversary?
-In the js paper, we assume syntactically well-formed, do we here assume that the capability for the adversary is in the value relation?
+In the JS paper, we assume syntactically well-formed, do we here assume that the capability for the adversary is in the value relation?
 If so in what world should this be true?
@@
 Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.
-If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$ \\
-\lau{Is $hs_{\var{frame}}$ supposed to be the rest of memory, so $h$ is something we can execute in?
-Is $\malloc$ allowed to allocate more memory, i.e., how does $\hs_{\var{footprint}}$ relate to $\hs_{\var{footprint}}'$?}\\
+If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$
+%\lau{Is $hs_{\var{frame}}$ supposed to be the rest of memory, so $h$ is something we can execute in?
+Is $\malloc$ allowed to allocate more memory, i.e., how does $\hs_{\var{footprint}}$ relate to $\hs_{\var{footprint}}'$?} A: The frame contains anything that has been allocated and the footprint contains the rest.
+That is the footprint contains mallocs internal state as well as the free memory.
+% Notice: the current worlds do not support malloc allocating memory as the invariants cannot evolve.
 that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$.
-Then $\malloc$ will succesfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}$\\
-\lau{doesn't this invariant allow the newly allocated memory to contain capabilities?}\\
- for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{?}{W'}$. 
-\lau{at what step should this hold? Just some step $n'<n$?}\\
-\lau{It should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.}\\
-\lau{How do we express that $\malloc$ can be ``trusted'' in the sense that it won't save the capab}
+Then $\malloc$ will successfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}^0$
+%\lau{doesn't this invariant allow the newly allocated memory to contain capabilities?} A: It does not capture this. One possiblity is to make new standard all zero invariant.
+ for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{n}{W'}$. 
+%\lau{at what step should this hold? Just some step $n'<n$?} A: try with n, if that does not work, then say something like: there exists j such that malloc terminates in j steps and use n-j.
+%\lau{It should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.} A: yes.
+If we won't abstract notion of this, the standard way to formulate it would be with Hoare triples.
+%\lau{How do we express that $\malloc$ can be ``trusted'' in the sense that it won't save any capabilities like one to the newly allocated memory or the return capability.} A:  it allows for read capabilities!
+To capture this, we need either non-interference or some kind of trace semantics!
+
+To \emph{invoke} $\malloc$ means that the capability in $\pcreg$ points to a $\jmp{\malloc}$ instruction (and that this address is within its capabilities).
+Further in the register $r_0$ there has to be a return capability, i.e., a capability of the form $(\perm,\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm$ is either, $\rwx$, $\exec$, or $\entry$.
+Lastly, the register $r_1$ has to the argument for $\malloc$ which is the desired length of the piece of memory to be allocated.
+
+To \emph{return} from $\malloc$ means that the register $\pcreg$ contains the capability $(\perm',\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm'$ is $exec$ if $\perm$ was $\exec$ and $\rwx$ otherwise.
+Further, register $r_1$ contains $(\rwx,\var{base},\var{end},\var{base})$ which is a capability for the allocated memory.
 \end{specification}
@@
 We also assume that there is a heap segment $\hs$ as described in the lemma.
-If we execute $\Phi$ until just before the jump to $\malloc$, then a number of steps has occured, say $0<i<n$, further location 0 on the heap has been overwritten with $-1$, the io capability has been saved to the designated address on the heap, and the $\rio$ register has been overwritten with 0.
+If we execute $\Phi$ until just before the jump to $\malloc$, then a number of steps has occurred, say $0<i<n$, further location 0 on the heap has been overwritten with $-1$, the io capability has been saved to the designated address on the heap, and the $\rio$ register has been overwritten with 0.
 
@@
 \end{align*}
-Then we need to argue that $\heapSat{\hs}{n-i}{W}$.
+Then we need to argue that $\heapSat[\hs]{n-i}{W}$.
 Given any address, in $\hs$ there is a region that covers that address, and the regions cover only the addresses in $\hs$.
 Now each of the heap segments have to be accepted by the respective regions.
-For the adversary, we assumed that it had no capabilities, so all the addresses contain intergers which i trivially in the value relation.
-The same goes for the heap segment with the ticket dispenser test program and the remaining addresses of the heap but addresses 0 \lau{need assumption about this.} The io address contains $-1$ at this point of the execution which mean that the heapsegment only consisting of address 0 is accepted in this world.
+For the adversary, we assumed that it had no capabilities, so all the addresses contain integers which i trivially in the value relation.
+The same goes for the heap segment with the ticket dispenser test program and the remaining addresses of the heap but addresses 0 \lau{need assumption about this.} The io address contains $-1$ at this point of the execution which mean that the heap segment only consisting of address 0 is accepted in this world.
 
@@
 
+A standard heap invariant that insures all values in the region are in the value  relation:
 \[
@@
                                      & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
+\end{align*}
+Another standard heap invariant that ensures all the words in the region are 0.
+\[
+  \iota_{\var{start},\var{end}}^0 \; : \; \StorePred
+\]
+\begin{align*}
+  \iota_{\var{start},\var{end}}^0(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
+                                     & \quad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
 \end{align*}
@@
 \begin{align*}
-  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: What should this be?
-  \stdkr(W) \defeq & \{ \npair{c} \mid \\
-                   & \quad \forall v \in \stdvr(W)\ldotp \\
-                   & \qquad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
-                   & \qquad \quad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
+  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: Check if this is still monotone
+  \stdkr(W) \defeq & \{ \npair{c} \mid c \in \stdvr(W) \land \\
+                   & \quad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
+                     \forall \heapSat{n}{W'} \ldotp \forall \reg \in \stdrr(W') \\
+                   & \qquad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
 \end{align*}
+We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.
+
+Normally, it would just be required for the return value to be in the value relation, so the continuation is invoked with this value.
+Here everything that is in the register file is available when the continuation is invoked, so we do not special case on $r_1$ (which is the register that in the calling convention we use should contain the return value).
+
 ``Expression'' relation:
@@
   For any $n \in \nats$, $W \in \Worlds$, and a word $\pc$, if  $\npair{\pc} \in \stdvr(W)$, then $\npair{\pc} \in \stder(W)$
-\end{lemma}
+\end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
+\lau{With the current definition of the expression relation, an entry pointer will never be in the expression relation, but I guess we do not want that restriction.
+Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
 

### 2016-06-17 capability-machine.tex
Finally add the nitpickings I promised.

--- 2016-06-16 capability-machine.tex
@@
 \lau{When continuation is invoked, it has to be in registers, but here we do not require it to be in the value relation, so it seems impossible to show some of the required conditions.}
+\dominique{$n'$ is unbound in the definition of $\stdkr$. }
+\dominique{$v \in \stdvr(W)$ should be $v \in \stdvr(W')$ in the definition of $\stdkr$? }
+\dominique{$\observations(W)$ should be $\observations(W')$ in the definition of $\stdkr$? }
 ``Expression'' relation:

### 2016-06-17 capability-machine.tex
added some comments/questions and updated a couple of definitions.

--- 2016-06-17 capability-machine.tex
@@
 \newcommand{\rio}{r_{io}}
+\newcommand{\proglen}{\var{pl}}
+\newcommand{\adv}{\var{adv}}
 \newcommand{\mrlen}{5cm}
@@
 \lau{maybe mention that malloc is also on the heap}
- \item[the heap segment $\boldsymbol{ \hs}$] contains all allocated addresses in the sense that any capability in $\Phi$ is a capability for a piece of memory in $\hs$,
- \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\init,\init+\var{proglen}-1,\init)$, which is the only capability that grants any access to this part of the heap,
+ \item[$\boldsymbol \memheap$] contains the malloc code starting at label $\malloc$.
+ \item[Nothing else is allocated in $\boldsymbol \memheap$] and all the above are disjoint.
+ \item[$\boldsymbol{\codelabel{adv}}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
+ \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\init,\init+\proglen-1,\init)$, which is the only capability that grants any access to this part of the heap,
  \item[$\boldsymbol{ \memreg(\rio)}$] is $(\readwrite,0,0,0)$, which also is the only capability that grants access to this part of the heap, 
@@
 The one with just $\iota_{\var{adv},\var{adv end}}$?}
-The variable $\var{proglen}$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
+The variable $\proglen$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
 
@@
 \item $\memreg(\rio) = (\readwrite,0,0,0)$ 
-\item $\memreg(\pcreg) = (\rwx,\init, \init + \var{proglen} - 1, \init)$
+\item $\memreg(\pcreg) = (\rwx,\init, \init + \proglen - 1, \init)$
 \item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
@@
   W(\codelabel{adv}) & = \iota_{\var{adv},\var{advend}} \\
-  W(\init)           & = \iota_{\init,\init+\var{proglen} - 1} \\
+  W(\init)           & = \iota_{\init,\init+\proglen - 1} \\
   W(\rio)            & = \lambda W\ldotp \{(n,\hs) \mid \hs(0) = -1 \text{ or } \\
@@
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  For any $n \in \nats$, $W \in \Worlds$, and a word $\pc$, if  $\npair{\pc} \in \stdvr(W)$, then $\npair{\pc} \in \stder(W)$
+  For any $n \in \nats$, $W \in \Worlds$, $\perm in \{\rwx, \exec, \entry\}$, and addresse $\var{base}$ and $\var{end}$, if $\writeCond(n,W,var{base},\var{end})$, then $\npair{(\perm, \var{base}, \var{end}, \var{base})} \in \stder(W)$.
 \end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
@@
 Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
+\lau{In the ticket dispenser lemma, we need to make some assumptions about the adversary to be able to use the FTLR.
+What should these assumptions be?
+So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition fromt his.}
 

### 2016-06-17 capability-machine.tex

@@
 It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result.
-As we will see in the semantics, it is assumed that the initial process starts with a read-write capability for this location.
+As we will see in the semantics, it is assumed that the inital process starts with a readwrite capability for this location.
 \begin{align*}
@@
 % Set io to -1
-\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialise io to $-1$} \\
+\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
 % Store io capability just after jump to adv
@@
 % Save counter
-           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialise counter to 0} \\
+           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialize counter to 0} \\
 % Change pointer to point to start of ticket dispenser code
@@
 % retrieve io capability
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capability} \\
+           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
            & \lea{r_0}{-2} & & \\
@@
 Here $\size$ is 9.
-The variables $\counter$ and $\iocap$ are respectively the offsets to the addresses reserved for the counter and io capabilities.
+The variables $\counter$ and $\iocap$ are resepctively the offsets to the addresses reserved for the counter and io capabilities.
 $i_1 \dots i_7$ refers to the instructions in the following ticket dispenser:
@@
 
+\dominique{$n'$ is unbound in the definition of $\stdkr$. }
+\dominique{$v \in \stdvr(W)$ should be $v \in \stdvr(W')$ in the definition of $\stdkr$? }
+\dominique{$\observations(W)$ should be $\observations(W')$ in the definition of $\stdkr$? }
 ``Expression'' relation:

### 2016-07-04 capability-machine.tex
typo
added response to Dominique comment
Remove comments that have been fixed by lau.
First attempt at the tdtp proof

--- 2016-06-17 capability-machine.tex
@@
 It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result.
-As we will see in the semantics, it is assumed that the initial process starts with a read-write capability for this location.
+As we will see in the semantics, it is assumed that the initial process starts with a readwrite capability for this location.
 \begin{align*}
@@
                            \end{cases} \\
-  \updatePcPerm{\perm,\start,\addrend,\addr} &=
+  \updatePcPerm{(\perm,\start,\addrend,\addr)} &=
                                      \begin{cases}
@@
                                              \begin{cases}
-                                               \updateReg{\pcreg}{\updatePcPerm{c}} &
+                                               \updateReg{\pcreg}{\updatePcPerm{\var{c}}} &
                                                \arraycolsep=0pt
@@
 \label{sec:tick-disp}
+\newcommand{\hsfoot}{\hs_\var{footprint}}
+\newcommand{\hsframe}{\hs_\var{frame}}
 \newcommand{\size}{\var{size}}
 \newcommand{\rio}{r_{io}}
-\newcommand{\proglen}{\var{pl}}
-\newcommand{\adv}{\var{adv}}
+\newcommand{\adv}{\codelabel{adv}}
+\newcommand{\advb}{\var{adv_{base}}}
+\newcommand{\adve}{\var{adv_{end}}}
+\newcommand{\initb}{\var{init}_{base}}
+\newcommand{\inite}{\var{init}_{end}}
 \newcommand{\mrlen}{5cm}
@@
 \newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
-Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\codelabel{adv}$.
+Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\advb$.
 Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).
-We assume that entry capabilities for $\codelabel{adv}$ and $\malloc$ is ambiently available.
+We assume that entry capabilities for $\adv$ and $\malloc$ is ambiently available.
 The following is a test program for a ticket dispenser:
@@
 % Set io to -1
-\init:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialise io to $-1$} \\
+\initb:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
 % Store io capability just after jump to adv
@@
 % jump to adv
-           & \jmp{\codelabel{adv}} & \bracket{1} & \annotate{1}{Jump to the adversary} \\
+           & \jmp{\adv} & \bracket{1} & \annotate{1}{Jump to the adversary} \\
            & 0 & \bracket{1} & \annotate{1}{Address reserved for io cap.} \\
@@
            & \store{\refheap{\rio}}{\refreg{r_0}} & & \\
-           & \halt
+\inite :   & \halt
   \end{array}
@@
  \begin{description}
- \item[$\boldsymbol \memheap$] contains the ticket dispenser test program starting at the address labelled $\init$, 
- \item[$\boldsymbol \memheap$] contains the adversary code starting at label $\codelabel{adv}$.
+ \item[$\boldsymbol \memheap$] contains the ticket dispenser test program starting at the address labelled $\initb$, 
+ \item[$\boldsymbol \memheap$] contains the adversary code starting at label $\advb$.
 The adversary is assumed to have no capabilities on the heap, but it has access to the ambient $\malloc$ capability!
 \lau{maybe mention that malloc is also on the heap}
- \item[$\boldsymbol \memheap$] contains the malloc code starting at label $\malloc$.
+ \item[$\boldsymbol \memheap$] contains the malloc code.
  \item[Nothing else is allocated in $\boldsymbol \memheap$] and all the above are disjoint.
- \item[$\boldsymbol{\codelabel{adv}}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
- \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\init,\init+\proglen-1,\init)$, which is the only capability that grants any access to this part of the heap,
+ \item[$\boldsymbol{\adv}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
+ \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\initb,\inite,\initb)$, which is the only capability that grants any access to this part of the heap,
  \item[$\boldsymbol{ \memreg(\rio)}$] is $(\readwrite,0,0,0)$, which also is the only capability that grants access to this part of the heap, 
@@
 \end{lemma}
-\lau{in the above lemma, what assumptions should we make on the adversary?
-In the JS paper, we assume syntactically well-formed, do we here assume that the capability for the adversary is in the value relation?
-If so in what world should this be true?
-The one with just $\iota_{\var{adv},\var{adv end}}$?}
-The variable $\proglen$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
+The in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
 
@@
 Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.
-If we invoke $\malloc$ in a memory $h = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}$
-%\lau{Is $hs_{\var{frame}}$ supposed to be the rest of memory, so $h$ is something we can execute in?
-Is $\malloc$ allowed to allocate more memory, i.e., how does $\hs_{\var{footprint}}$ relate to $\hs_{\var{footprint}}'$?} A: The frame contains anything that has been allocated and the footprint contains the rest.
+If we invoke $\malloc$ in a memory $h = \hsframe \uplus \hsfoot$
+%\lau{Is $\hsframe$ supposed to be the rest of memory, so $h$ is something we can execute in?
+Is $\malloc$ allowed to allocate more memory, i.e., how does $\hsfoot$ relate to $\hsfoot'$?} A: The frame contains anything that has been allocated and the footprint contains the rest.
 That is the footprint contains mallocs internal state as well as the free memory.
 % Notice: the current worlds do not support malloc allocating memory as the invariants cannot evolve.
-that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hs_{\var{footprint}}]{n}{W_{\malloc}}$.
-Then $\malloc$ will successfully terminate and return a result in a heap $h' = \hs_{\var{frame}} \uplus \hs_{\var{footprint}}'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}^0$
+that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hsfoot]{n}{W_{\malloc}}$.
+Then $\malloc$ will successfully terminate and return a result in a heap $h' = \hsframe \uplus \hsfoot'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}^0$
 %\lau{doesn't this invariant allow the newly allocated memory to contain capabilities?} A: It does not capture this. One possiblity is to make new standard all zero invariant.
- for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hs_{\var{footprint}}']{n}{W'}$. 
+ for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hsfoot']{n}{W'}$. 
 %\lau{at what step should this hold? Just some step $n'<n$?} A: try with n, if that does not work, then say something like: there exists j such that malloc terminates in j steps and use n-j.
@@
  \begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
-Let configuration $\Phi$ be given.
-Assume that the ticket dispenser test code resides in memory starting at address $\init$ along with the adversary code at address $\codelabel{adv}$.
-Also assume that capabilities for $\codelabel{adv}$ and $\malloc$ are ambiently available.
-Further assume that the register file contains the following:
+Let configuration $\Phi$ be given where the heap contains:
+\begin{itemize}
+\item $\memheap$ contains the ticket dispenser test program (tdtp) starting at address $\initb$.
+\item $\memheap$ contains the adversary code starting at address $\advb$ and ending at address $\adve$. The adversary code contains no capabilities.
+\item $\memheap$ contains the malloc code.
+\item (and so on) %TODO write the remaining assumptions
+\end{itemize}
+and the register file contains:
 \begin{enumerate}
 \item $\memreg(\rio) = (\readwrite,0,0,0)$ 
-\item $\memreg(\pcreg) = (\rwx,\init, \init + \proglen - 1, \init)$
+\item $\memreg(\pcreg) = (\rwx,\initb, \inite, \initb)$
 \item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
 \end{enumerate}
-Further assume $\Phi \rightarrow^* (\halted,\Phi')$.
-As the evaluation terminates, this must happen in a finite number of steps, say this number is $n$.
-We also assume that there is a heap segment $\hs$ as described in the lemma.
-If we execute $\Phi$ until just before the jump to $\malloc$, then a number of steps has occurred, say $0<i<n$, further location 0 on the heap has been overwritten with $-1$, the io capability has been saved to the designated address on the heap, and the $\rio$ register has been overwritten with 0.
 
-Now we need to jump to $\malloc$, so we need to make sure that the precondition in the specification is satisfied.
-To do this, we first need a world that models the part of the heap that is currently in use which is $\hs$.
-We pick the world $W$ to be:
-\begin{align*}
-  W(\codelabel{adv}) & = \iota_{\var{adv},\var{advend}} \\
-  W(\init)           & = \iota_{\init,\init+\proglen - 1} \\
-  W(\rio)            & = \lambda W\ldotp \{(n,\hs) \mid \hs(0) = -1 \text{ or } \\
-                     &   \hphantom{ = \lambda W\ldotp \{(n,\hs) \mid }\:\; \hs(0) \text{ even and positive} \}
-\end{align*}
-for all other addresses $a$ that are not captured in the above but are part of $hs$, we define a new region $r$ in $W$ where
-\begin{align*}
-  W(r) & = \iota_{a,a}
-\end{align*}
-Then we need to argue that $\heapSat[\hs]{n-i}{W}$.
-Given any address, in $\hs$ there is a region that covers that address, and the regions cover only the addresses in $\hs$.
-Now each of the heap segments have to be accepted by the respective regions.
-For the adversary, we assumed that it had no capabilities, so all the addresses contain integers which i trivially in the value relation.
-The same goes for the heap segment with the ticket dispenser test program and the remaining addresses of the heap but addresses 0 \lau{need assumption about this.} The io address contains $-1$ at this point of the execution which mean that the heap segment only consisting of address 0 is accepted in this world.
+If we start running in $\Phi$ until $\memreg(\pcreg)$ points to $\jmp{\malloc}$, then the configuration is the same as $\Phi$ except:
+\begin{itemize}
+\item $\memheap(0)$ contains $-1$
+\item the io capabililty is saved to the designated address in the tdtp
+\item $r_1$ contains the size of the ticket dispenser program (9)
+\item $r_0$ contains the return capability $(\rwx,\initb,\inite, \initb+\var{mallocOffset}+1)$
+\end{itemize}
+Call this new heap $\Phi'$
 
-The remaining conditions in the specification are also satisfied, so there exists a configuration $\Phi'$
+To be able to say something about the jump to $\malloc$, we use its specification.
+Take $\hsfoot$ to be the part of the heap $\malloc$ resides in and everything that has not been allocated according to the assumptions.
+The $\hsframe$ to be everything else.
+We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.
+According to the specification of $\malloc$, we get $\Phi' \step^* \Phi''$ where
+\begin{itemize}
+\item $\memheap[\Phi'] = \hsframe \uplus \hsfoot'$
+\item $\memreg[\Phi'](r_1) = (\rwx,\var{base},\var{end},\var{base})$ for some addresses $\var{base}$ and $\var{end}$
+\item $\memreg[\Phi'](\pcreg)$
+\end{itemize}
+Further for $W'$ with exactly two regions $\iota_\malloc$ and $\iota_{\var{base},\var{end}}^0$ we have that the new footprint satisfies this world, i.e., $\heapSat[\hsfoot']{n}{W'}$.
+
+If we continue running from $\Phi'$ until just after the point we have executed $\jmp{\adv}$, then  $\pcreg$ points to an execute capability for the adversary, and  we reach a configuration $\Phi''$ which is the same as $\Phi$ except:
+\begin{itemize}
+\item $\memheap[\Phi''](0)$ contains $-1$
+\item capabilities for \emph{the counter} and io are saved to the designated addresses in the tdtp
+\item the ticket dispenser code resides on the heap starting at the address $\var{base}$ (the code is as seen in Figure~\ref{tab:tick-disp-heap})
+\item $\memreg[\Phi''](\pcreg) = (\exec,\advb,\adve,\advb)$
+\item $\memreg[\Phi''](r_0) = (\entry,\initb,\inite,\initb+\var{ret})$ for an appropriate $\var{ret}$ such that the return pointer points to the instruction $\move{\refreg{r_0}{\pcreg}}$, just after the memory designated for capabilities.
+\item $\memreg[\Phi''](r_1) = (\entry,\var{base},\var{end},\var{base})$
+\item all other registers contain 0
+\end{itemize}
+We now want to use the FTLR. To do so, we need to define a world that represents our current heap. We pick the following world:
+\begin{itemize}
+\item $\iota_{\var{io}}$ the region governs the heap segment that only has address $0$. The value of address $0$ should be $-1$ or $\geq 0$ and even.  % The region that governs that output address
+\item $\iota_{\var{td}}$ this region governs heap segments with the domain $[\var{base},\var{end}]$ for the addresses $[\var{base},\var{end} - 1]$ they are the same as in $\Phi''$.
+$\hs(\var{base})$ is positive and even.
+\item $\iota_{\var{tdtp}}$ this region governs the ticket dispenser test programs.
+The domain of a heap segment in this region is that of the ticket dispenser test program and the contents of the heap segment should be the same as it is in $\Phi''$.
+\item $\iota_{\advb,\adve}$
+\item $\iota_{\malloc}$
+\end{itemize}
+To use the FTLR, we have to argue that the read condition is satisfied for the region, the adversary is in. This is easily shown, as we have chosen the region for the adversary to be the standard one.
+
+so we wan conclude that $\npair{(\exec,\advb,\adve,\advb}) \in \stder(W'')$ where $W''$ is the world with the following regions:
+Assuming we can show the following three things:
+\begin{enumerate}
+\item $\memreg[\Phi''] \in \stdrr(W'')$
+\item $\npair{c = (\exec,\initb,\inite,\init+\var{ret})}] \in \stdkr(W'')$
+\item $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$ where the we have limited the heap to the segments governed by the regions in $W''$.
+\end{enumerate}
+then it follows from the fact that the adversary program is in the expression predicate that $\npair{\Phi^{(3)} = (\updateReg[\Phi'']{r_0}{c}[\plainproj{reg}.r_1 \mapsto (\exec,\advb,\adve,\advb)].\plainproj{reg},\memheap[\Phi''])})] \in \observations(W'')$.
+From this and the fact that we know the execution terminates, we know that $\Phi^{(3)} \step^k (\halted,h)$ for some heap $h$ and $k \leq n$.
+From this we can conclude that the heap segment that only consists of address 0 is accepted by region $\rio$ which means that either $h(0)$ is $-1$ or positive and even which is exactly what the lemma says.
+
+It remains to that 1-3 holds.
+For 1, all registers but $r_0$ and $r_1$ are easy as they contain $0$ which is always in the value relation.
+For $r_0$ and $r_1$, we do need to do some reasoning.
+Both capabilities have one things in common, namely that the underlying programs do not use the arguments.
+
+$\memreg[\Phi''](r_0)$ contains the capability $(\entry,\initb,\inite,\initb+\var{ret})$ and we need to show that it is in $\stdvr(W'')$.
+Let $n' < n$ and $W^{(3)} \future W''$ be given and show $\npair[n']{(\exec,\initb,\inite,\initb+\var{ret})} \in \stder(W^{(3)})$.
+Let $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
+We do not really care about the register file and the continuation as the ticket dispenser test program does not use them.
+As $\heap$ satisfies $W^{(3)}$ and parts of the heap must satisfy the invariants $\iota_{\var{tdtp}}$ and $\iota_{\var{td}}$ which means the ticket dispenser program and the ticket dispenser test program is unaltered.
+It also means that the counter in the ticket dispenser is $\geq 0$ and even.
+We need to show $\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\initb,\inite,\initb+\var{ret}), \heap \uplus \heap_f)} \in \observations(W^{(3)})$.
+If $n'$ is sufficiently small, then we run out of steps before the program halts in which case the configuration is trivially in the $\observations$ predicate.
+If the execution halts successfully, then the ticket dispenser test program will have moved the value of the counter in the ticket dispenser to address 0, the designated io address.
+As this is a numerical value, it will trivially be in the value relation.
+Further, the island that governs address 0 i $\iota_{\var{io}}$, so we have to argue that the value is $\geq 0$ and even.
+The value came from the ticket dispenser which is governed by the region $\iota_{\var{td}}$ so it is indeed $\geq 0$ and even.
+
+In register $r_1$ we have the capability $(\entry,\var{base},\var{end},\var{base})$ and we need to show $\npair[n]{(\exec,\var{base},\var{end},\var{base})} \in \stder(W'')$.
+To do this end, we go through the same motion as for $r_0$:  Let $n' < n$, $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
+Show $\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\var{base},\var{end},\var{base}), \heap \uplus \heap_f)} \in \observations(W^{(3)})$.
+Again the heap satisfaction makes sure that the ticket dispenser program is as we expect, so we can execute it.
+If we run out of steps before we reach $\jmp{\refreg{r_0}}$, then it is in the observation predicate.
+Otherwise, we reach a new configuration $\Phi^{(3)}$ where the registers are the same as in $\reg$ except for $r_0$, $r_2$, and $\pcreg$ which will be $c$, $0$, and $\updatePcPerm(c)$ respectively.
+$0$ is trivially in the value relation and $c$ is in the value relation as it is in the continuation relation, so $\reg$ is in the register-file relation.
+If we split the heap part of $\Phi^{(3)}$ into the same two parts as we had in the initial assumption, then we need to argue that the non-frame part still satisfies the world.
+The only change made to the heap was updating the counter, so there are heap segments that satisfy all regions but $\iota_{\var{td}}$ as they are unchanged.
+The ticket dispenser satisfies $\iota_{\var{td}}$ as everything but the counter remains unchanged, so we just need to argue that the counter value is still $\geq 0$ and even.
+This value came from the previous counter value which was in a heap segment that satisfied $\iota_{\var{td}}$, so it was positive and even.
+We then added two and stored it which again gives a positive and even value.
+So the ticket dispenser part of the heap satisfies $\iota_{\var{td}}$.
+Now we have all the necessary conditions satisfied to be able to use the fact that $c$ is in the continuation relation to conclude
+\lau{did I mess up the steps here?}
+$\npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\updatePcPerm{c}}, \heap' \uplus \heap_f)}$ (where $\heap'$ is the new heap).
+From this we get that continuing from the configuration we end up with no more steps, failing, or in a halting state that satisfy the world.
+
+%condition 2
+For condition 2, we need to show $\npair{c = (\entry, \initb, \inite,\initb+\var{ret})}] \in \stdkr(W'')$.
+To this end, let $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
+We then need to argue
+\[
+\npair{(\reg\update{\pcreg}{\updatePcPerm{c}},\heap \uplus \heap_f)} \in \observations(W^{(3)})
+\]
+As the heap satisfy a future world of $W''$, we know that the ticket dispenser test program remains unaltered.
+If we run it (without running out of steps), then it loads the couner, stores it to address $0$, and halts.
+From the heap satisfaction, we get that the counter is positive and even, so the value we store to address 0 is positive and even.
+The value is trivially in the value relation as it is an integer.
+Further the region that governs address 0 is $\iota_{\var{io}}$ which requires the value there to be positive and even which it is.
+So in this case the configuration is in the observation predicate.
+If we run out of steps, then it is also in the observation relation.
+
+%condition 3
+Finally, we need to show $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$.
+The domains of the heap segments are assumed to be disjoined and we have limited the heap to only be the once concidered in the regions, so the firs condition of heap satisfaction is okay.
+Let us take the regions one at a time and shortly argue that the corresponding heap segment satisfies this region:
+\begin{itemize}
+\item $\iota_{\var{io}}$ governs address 0, which i initial $-1$. This region requires it to be $-1$ or $\geq 0$ and even, so it is initially satisfied.
+\item $\iota_{\var{tdtp}}$, this region was constructed so the part of the heap it governs remains unchanged, so it trivially satisfy this region.
+\item $\iota_{\var{td}}$, apart from the address with the counter, this region is also constructed so this heap segment does not change.
+With respect to the counter, it requires it to be $\geq 0$ and even.
+The counter starts as 0, so also this cell satisfies the region.
+\item $\iota_{\advb,\adve}$ the adversary code is assumed to be instructions only which are trivially satisfied by the standard island.
+\item $\iota_{\malloc}$, from the specification of $\malloc$ we got that the resulting heap after the allocation satisfied a world with $\iota_{\malloc}$.
+We use the same region and we have not changed that part of the heap, so it still satifies this region.
+\end{itemize}
+
 \end{proof}
@@
 \begin{align*}
-  \iota_{\var{start},\var{end}}(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
+  \iota_{\var{start},\var{end}}(W) \defeq  \{ \npair{\heapseg} \mid & \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
@@
 \begin{align*}
-  \iota_{\var{start},\var{end}}^0(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
-                                     & \quad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
+  \iota_{\var{start},\var{end}}^0(W) \defeq \{ \npair{\heapseg} \mid & \dom(\heapseg) = [\var{start},\var{end}] \land \\
+                                                                & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
 \end{align*}
@@
   \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                          & \; \quad \npair[n']{(\rwx,\start,\addrend,\addr)} \in \stder(W') \\
+                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
 \end{align*}
@@
                            & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
-                           & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
+                           & \quad (\exists i < n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                            & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap')  \\
@@
 \end{align*}
+\lau{Is the last condition sufficient? If any of the other two conditions are true, then the last becomes trivially true as well.}\\
 The ``good observations'' are as follows
@@
                    & \quad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat{n}{W'} \ldotp \forall \reg \in \stdrr(W') \\
-                   & \qquad (n,reg\update{r_0}{v}\update{\pcreg}{\updatePcPerm{c}}, \heap) \in \observations(W) \}
+                     \forall \heapSat{n}{W'}, \heap_f \ldotp \forall \reg \in \stdrr(W') \\
+                   & \qquad \npair{(reg\update{\pcreg}{\updatePcPerm{\var{c}}}, \heap \uplus \heap_f} \in \observations(W') \}
 \end{align*}
@@
                    & \qquad \quad  \forall \heapSat{n}{W}, \heap_f \ldotp \\
-                   & \qquad \qquad (n,\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f) \in \observations(W) \}
+                   & \qquad \qquad \npair{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f)} \in \observations(W) \}
 \end{align*}
@@
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  For any $n \in \nats$, $W \in \Worlds$, $\perm in \{\rwx, \exec, \entry\}$, and addresse $\var{base}$ and $\var{end}$, if $\writeCond(n,W,var{base},\var{end})$, then $\npair{(\perm, \var{base}, \var{end}, \var{base})} \in \stder(W)$.
+  For any $n \in \nats$, $W \in \Worlds$, and addresses $\var{base}$ and $\var{end}$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, \var{base})} \in \stder(W)$.
 \end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
@@
 What should these assumptions be?
-So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition fromt his.}
+So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition from this.}
 

### 2016-07-05 capability-machine.tex
Fix a few typos and add some remarks.

@@
 Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).
-We assume that entry capabilities for $\adv$ and $\malloc$ is ambiently available.
+We assume that entry capabilities for $\adv$ and $\malloc$ is ambiently available\dominique{what does this mean?}.
 The following is a test program for a ticket dispenser:
@@
 The adversary could also choose never to return to the return address, which we set up for him.
-\begin{lemma}[Ticket dispenser test]
+\begin{lemma}[Ticket dispenser test]\DD{I would prefer more precise, symbolic statements of assumptions and less prose.}
 \label{lem:tckt-disp}
@@
 If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
-\begin{specification}[Malloc]
+\begin{specification}[Malloc]\dominique{can you state this more systematically, with less prose, more explicit assumptions and conclusion?}
 Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.
@@
 \item $\memheap(0)$ contains $-1$
-\item the io capabililty is saved to the designated address in the tdtp
+\item the io capability is saved to the designated address in the tdtp
 \item $r_1$ contains the size of the ticket dispenser program (9)
@@
 The $\hsframe$ to be everything else.
-We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.
-According to the specification of $\malloc$, we get $\Phi' \step^* \Phi''$ where
+We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.\dominique{explicit assumption about this needed?} According to the specification of $\malloc$, we get $\Phi' \step^* \Phi''$ where
 \begin{itemize}
@@
 
-so we wan conclude that $\npair{(\exec,\advb,\adve,\advb}) \in \stder(W'')$ where $W''$ is the world with the following regions:
+So we can conclude that $\npair{(\exec,\advb,\adve,\advb}) \in \stder(W'')$ where $W''$ is the world with the following regions:
 Assuming we can show the following three things:
@@
 As this is a numerical value, it will trivially be in the value relation.
-Further, the island that governs address 0 i $\iota_{\var{io}}$, so we have to argue that the value is $\geq 0$ and even.
+Further, the island that governs address 0 is $\iota_{\var{io}}$, so we have to argue that the value is $\geq 0$ and even.
 The value came from the ticket dispenser which is governed by the region $\iota_{\var{td}}$ so it is indeed $\geq 0$ and even.
@@
 If we run out of steps before we reach $\jmp{\refreg{r_0}}$, then it is in the observation predicate.
-Otherwise, we reach a new configuration $\Phi^{(3)}$ where the registers are the same as in $\reg$ except for $r_0$, $r_2$, and $\pcreg$ which will be $c$, $0$, and $\updatePcPerm(c)$ respectively.
+Otherwise, we reach a new configuration $\Phi^{(3)}$ where the registers are the same as in $\reg$ except for $r_0$, $r_2$, and $\pcreg$ which will be $c$, $0$, and $\updatePcPerm{c}$ respectively.
 $0$ is trivially in the value relation and $c$ is in the value relation as it is in the continuation relation, so $\reg$ is in the register-file relation.
@@
 %condition 2
+\dominique{this duplicates the argument about $(\entry,\initb,\inite,\initb+\var{ret}) \in \stdvr(W'')$. Find a way to share this?}
 For condition 2, we need to show $\npair{c = (\entry, \initb, \inite,\initb+\var{ret})}] \in \stdkr(W'')$.
@@
 Finally, we need to show $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$.
-The domains of the heap segments are assumed to be disjoined and we have limited the heap to only be the once concidered in the regions, so the firs condition of heap satisfaction is okay.
+The domains of the heap segments are assumed to be disjoint and we have limited the heap to only be the parts considered in the regions, so the first condition of heap satisfaction is okay.
 Let us take the regions one at a time and shortly argue that the corresponding heap segment satisfies this region:

### 2016-07-05 capability-machine.tex
Add another comment.
Oops, latex fix.

@@
 The adversary could also choose never to return to the return address, which we set up for him.
-\begin{lemma}[Ticket dispenser test]\DD{I would prefer more precise, symbolic statements of assumptions and less prose.}
+\begin{lemma}[Ticket dispenser test]\dominique{I would prefer more precise, symbolic statements of assumptions and less prose.}
 \label{lem:tckt-disp}
@@
 So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition from this.}
+\dominique{Something I find interesting here, is that the FTLR can be interpreted to say that x permissions have no real security value: if you have an arbitrary r or rw capability that is valid, then the FTLR says that it remains valid if we change it to rx or rwx (respectively).}
 

### 2016-07-05 capability-machine.tex
Made some changes to the proof of lemma 1 to enhance readability and correct mistakes

--- 2016-07-04 capability-machine.tex
@@
 \newcommand{\mrlen}{5cm}
+\newcommand{\retm}{\var{ret}_{\malloc}}
+\newcommand{\reta}{\var{ret}_{\adv}}
 \newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
@@
 % Save ticket dispenser code
-           & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
+\retm      & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
            & \lea{\refreg{r_1}}{1} & & \\
@@
 % retrieve io capability
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
+\reta      & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
            & \lea{r_0}{-2} & & \\
@@
 In particular, the adversary can execute a $\halt$ instruction that will cause the machine to halt (successfully).
-The adversary could also choose never to return to the return address, which we set up for him.
+The adversary could also choose never to return to the return address that we set up for him.
 \begin{lemma}[Ticket dispenser test]
@@
  \item[$\boldsymbol \memheap$] contains the malloc code.
- \item[Nothing else is allocated in $\boldsymbol \memheap$] and all the above are disjoint.
+ \item[Nothing else is allocated in $\boldsymbol \memheap$] and all the above programs are disjoint.
  \item[$\boldsymbol{\adv}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
@@
 \end{lemma}
-The in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
+In the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
 
 The ticket dispenser test program uses $\malloc$.
-To be able to reason about programs that use $\malloc$, we assume that it obeys the following specification.
+To be able to reason about programs that use $\malloc$, we assume that it follows the following specification.
 If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
@@
 Further in the register $r_0$ there has to be a return capability, i.e., a capability of the form $(\perm,\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm$ is either, $\rwx$, $\exec$, or $\entry$.
-Lastly, the register $r_1$ has to the argument for $\malloc$ which is the desired length of the piece of memory to be allocated.
+Lastly, the register $r_1$ has to contain the argument for $\malloc$ which is the desired length of the piece of memory to be allocated.
 
-To \emph{return} from $\malloc$ means that the register $\pcreg$ contains the capability $(\perm',\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm'$ is $exec$ if $\perm$ was $\exec$ and $\rwx$ otherwise.
+To \emph{return} from $\malloc$ means that the register $\pcreg$ contains the capability $(\perm',\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm'$ is $\rwx$ if $\perm$ was $\rwx$ and $\exec$ otherwise.
 Further, register $r_1$ contains $(\rwx,\var{base},\var{end},\var{base})$ which is a capability for the allocated memory.
@@
  \begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
-Let configuration $\Phi$ be given where the heap contains:
+Let configuration $\Phi$ be given where the heap is shaped as follows:
 \begin{itemize}
@@
 \item $r_1$ contains the size of the ticket dispenser program (9)
-\item $r_0$ contains the return capability $(\rwx,\initb,\inite, \initb+\var{mallocOffset}+1)$
+\item $r_0$ contains the return capability $(\rwx,\initb,\inite, \retm)$
 \end{itemize}
@@
 Take $\hsfoot$ to be the part of the heap $\malloc$ resides in and everything that has not been allocated according to the assumptions.
-The $\hsframe$ to be everything else.
+Take $\hsframe$ to be everything else.
 We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.
@@
 \item $\memreg[\Phi'](r_1) = (\rwx,\var{base},\var{end},\var{base})$ for some addresses $\var{base}$ and $\var{end}$
-\item $\memreg[\Phi'](\pcreg)$
+\item $\memreg[\Phi'](\pcreg) = (\rwx,\initb,\inite,\retm)$
 \end{itemize}
@@
 \item $\memreg[\Phi''](\pcreg) = (\exec,\advb,\adve,\advb)$
-\item $\memreg[\Phi''](r_0) = (\entry,\initb,\inite,\initb+\var{ret})$ for an appropriate $\var{ret}$ such that the return pointer points to the instruction $\move{\refreg{r_0}{\pcreg}}$, just after the memory designated for capabilities.
+\item $\memreg[\Phi''](r_0) = (\entry,\initb,\inite,\reta)$
 \item $\memreg[\Phi''](r_1) = (\entry,\var{base},\var{end},\var{base})$
@@
 \end{itemize}
-To use the FTLR, we have to argue that the read condition is satisfied for the region, the adversary is in. This is easily shown, as we have chosen the region for the adversary to be the standard one.
+To use the FTLR, we have to argue that the read condition is satisfied for the region, the adversary is in.
+This is easily shown, as we have chosen the region for the adversary to be the standard one, so with $\advb$ and $\adve$ as witnesses, we need to show $\iota{\advb,\adve} \subseteq \iota{\advb,\adve}$.
 
@@
 \item $\memreg[\Phi''] \in \stdrr(W'')$
-\item $\npair{c = (\exec,\initb,\inite,\init+\var{ret})}] \in \stdkr(W'')$
+\item $\npair{c = (\exec,\initb,\inite,\reta)}] \in \stdkr(W'')$
 \item $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$ where the we have limited the heap to the segments governed by the regions in $W''$.
 \end{enumerate}
-then it follows from the fact that the adversary program is in the expression predicate that $\npair{\Phi^{(3)} = (\updateReg[\Phi'']{r_0}{c}[\plainproj{reg}.r_1 \mapsto (\exec,\advb,\adve,\advb)].\plainproj{reg},\memheap[\Phi''])})] \in \observations(W'')$.
+then it follows from the fact that the adversary program is in the expression predicate that
+\[
+\npair{\Phi^{(3)} = (\updateReg[\Phi'']{r_0}{c}[\plainproj{reg}.r_1 \mapsto (\exec,\advb,\adve,\advb)].\plainproj{reg},\memheap[\Phi''])} \in \observations(W'').
+\]
 From this and the fact that we know the execution terminates, we know that $\Phi^{(3)} \step^k (\halted,h)$ for some heap $h$ and $k \leq n$.
-From this we can conclude that the heap segment that only consists of address 0 is accepted by region $\rio$ which means that either $h(0)$ is $-1$ or positive and even which is exactly what the lemma says.
+From this we can conclude that the heap segment that only consists of address 0 is accepted by region $\iota_{\var{io}}$ which means that either $h(0)$ is $-1$ or positive and even which is exactly what the lemma says.
 
-It remains to that 1-3 holds.
-For 1, all registers but $r_0$ and $r_1$ are easy as they contain $0$ which is always in the value relation.
+It remains to show that 1-3 holds.
+For 1, all registers but $r_0$ and $r_1$ are easy as they contain $0$ which is always in the value predicate.
 For $r_0$ and $r_1$, we do need to do some reasoning.
@@
 
-$\memreg[\Phi''](r_0)$ contains the capability $(\entry,\initb,\inite,\initb+\var{ret})$ and we need to show that it is in $\stdvr(W'')$.
-Let $n' < n$ and $W^{(3)} \future W''$ be given and show $\npair[n']{(\exec,\initb,\inite,\initb+\var{ret})} \in \stder(W^{(3)})$.
+$\memreg[\Phi''](r_0)$ contains the capability $(\entry,\initb,\inite,\reta)$ and we need to show that it is in $\stdvr(W'')$.
+Let $n' < n$ and $W^{(3)} \future W''$ be given and show $\npair[n']{(\exec,\initb,\inite,\reta)} \in \stder(W^{(3)})$.
 Let $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
 We do not really care about the register file and the continuation as the ticket dispenser test program does not use them.
-As $\heap$ satisfies $W^{(3)}$ and parts of the heap must satisfy the invariants $\iota_{\var{tdtp}}$ and $\iota_{\var{td}}$ which means the ticket dispenser program and the ticket dispenser test program is unaltered.
+As $\heap$ satisfies $W^{(3)}$, parts of the heap must satisfy the invariants $\iota_{\var{tdtp}}$ and $\iota_{\var{td}}$ which means the ticket dispenser program and the ticket dispenser test program is unaltered.
 It also means that the counter in the ticket dispenser is $\geq 0$ and even.
-We need to show $\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\initb,\inite,\initb+\var{ret}), \heap \uplus \heap_f)} \in \observations(W^{(3)})$.
+We need to show $\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\initb,\inite,\reta), \heap \uplus \heap_f)} \in \observations(W^{(3)})$.
 If $n'$ is sufficiently small, then we run out of steps before the program halts in which case the configuration is trivially in the $\observations$ predicate.
-If the execution halts successfully, then the ticket dispenser test program will have moved the value of the counter in the ticket dispenser to address 0, the designated io address.
-As this is a numerical value, it will trivially be in the value relation.
-Further, the island that governs address 0 i $\iota_{\var{io}}$, so we have to argue that the value is $\geq 0$ and even.
+If the execution halts successfully, then the ticket dispenser test program will have moved the value of the counter in the ticket dispenser to address 0, the designated I/O address.
+As this is a numerical value, it will trivially be in the value predicate.
+Further, the island that governs address 0 is $\iota_{\var{io}}$, so we have to argue that the value is $\geq 0$ and even.
 The value came from the ticket dispenser which is governed by the region $\iota_{\var{td}}$ so it is indeed $\geq 0$ and even.
@@
 In register $r_1$ we have the capability $(\entry,\var{base},\var{end},\var{base})$ and we need to show $\npair[n]{(\exec,\var{base},\var{end},\var{base})} \in \stder(W'')$.
-To do this end, we go through the same motion as for $r_0$:  Let $n' < n$, $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
-Show $\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\var{base},\var{end},\var{base}), \heap \uplus \heap_f)} \in \observations(W^{(3)})$.
+To this end, we go through the same motion as for $r_0$:  Let $n' < n$, $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
+Show
+\[
+\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\var{base},\var{end},\var{base}), \heap \uplus \heap_f)} \in \observations(W^{(3)}).
+\]
 Again the heap satisfaction makes sure that the ticket dispenser program is as we expect, so we can execute it.
 If we run out of steps before we reach $\jmp{\refreg{r_0}}$, then it is in the observation predicate.
-Otherwise, we reach a new configuration $\Phi^{(3)}$ where the registers are the same as in $\reg$ except for $r_0$, $r_2$, and $\pcreg$ which will be $c$, $0$, and $\updatePcPerm(c)$ respectively.
-$0$ is trivially in the value relation and $c$ is in the value relation as it is in the continuation relation, so $\reg$ is in the register-file relation.
+Otherwise, we reach a new configuration $\Phi^{(3)}$ where the registers are the same as in $\reg$ except for $r_0$, $r_1$, $r_2$, and $\pcreg$ which will be $c$, the previous counter value, $0$, and $\updatePcPerm(c)$ respectively.
+$0$ is trivially in the value predicate and $c$ is in the value predicate as it is in the continuation predicate, so $\reg$ is in the register-file predicate.
 If we split the heap part of $\Phi^{(3)}$ into the same two parts as we had in the initial assumption, then we need to argue that the non-frame part still satisfies the world.
-The only change made to the heap was updating the counter, so there are heap segments that satisfy all regions but $\iota_{\var{td}}$ as they are unchanged.
-The ticket dispenser satisfies $\iota_{\var{td}}$ as everything but the counter remains unchanged, so we just need to argue that the counter value is still $\geq 0$ and even.
-This value came from the previous counter value which was in a heap segment that satisfied $\iota_{\var{td}}$, so it was positive and even.
-We then added two and stored it which again gives a positive and even value.
-So the ticket dispenser part of the heap satisfies $\iota_{\var{td}}$.
-Now we have all the necessary conditions satisfied to be able to use the fact that $c$ is in the continuation relation to conclude
+The only change made to the heap was updating the counter.
+The heap segments that did not change still satisfy their regions.
+We have to argue that the heap segment $[\var{base},\var{end}]$ still satisfies the region $\iota_{\var{td}}$.
+The ticket dispenser code remains unchanged and the counter is $\geq 0$ and even as it was computed by adding two to a value which was $\geq 0$ and even, so the ticket dispenser heap segment satisfies $\iota_{\var{td}}$.
+Now we have all the necessary conditions satisfied to be able to use the fact that $c$ is in the continuation predicate to conclude
 \lau{did I mess up the steps here?}
-$\npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\updatePcPerm{c}}, \heap' \uplus \heap_f)}$ (where $\heap'$ is the new heap).
-From this we get that continuing from the configuration we end up with no more steps, failing, or in a halting state that satisfy the world.
+\[
+\npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\updatePcPerm{c}}, \heap' \uplus \heap_f)} \in \observations
+\]
+(where $\heap'$ is the new heap). From this we get that continuing from the configuration we end up with no more steps, failing, or in a halting state that satisfies the world.
 
 %condition 2
-For condition 2, we need to show $\npair{c = (\entry, \initb, \inite,\initb+\var{ret})}] \in \stdkr(W'')$.
-To this end, let $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
-We then need to argue
+For condition 2, we need to show 
 \[
-\npair{(\reg\update{\pcreg}{\updatePcPerm{c}},\heap \uplus \heap_f)} \in \observations(W^{(3)})
+\npair{c = (\entry, \initb, \inite,\reta)}] \in \stdkr(W'').
 \]
-As the heap satisfy a future world of $W''$, we know that the ticket dispenser test program remains unaltered.
-If we run it (without running out of steps), then it loads the couner, stores it to address $0$, and halts.
+To this end, let $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given. We then need to argue 
+\[
+\npair{(\reg\update{\pcreg}{\updatePcPerm{c}},\heap \uplus \heap_f)} \in \observations(W^{(3)}).
+\]
+As part of the heap satisfies a future world of $W''$, we know that the ticket dispenser test program remains unaltered.
+If we run it (and do not run out of steps), then it loads the couner, stores it to address $0$, and halts.
 From the heap satisfaction, we get that the counter is positive and even, so the value we store to address 0 is positive and even.
@@
 Finally, we need to show $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$.
-The domains of the heap segments are assumed to be disjoined and we have limited the heap to only be the once concidered in the regions, so the firs condition of heap satisfaction is okay.
-Let us take the regions one at a time and shortly argue that the corresponding heap segment satisfies this region:
+The domains of the heap segments are assumed to be disjoined, and we have limited the heap to only be the parts considered in the regions, so the first condition of heap satisfaction is okay.
+Let us take the regions one at a time and briefly argue that the corresponding heap segment satisfies the region:
 \begin{itemize}
 \item $\iota_{\var{io}}$ governs address 0, which i initial $-1$. This region requires it to be $-1$ or $\geq 0$ and even, so it is initially satisfied.
-\item $\iota_{\var{tdtp}}$, this region was constructed so the part of the heap it governs remains unchanged, so it trivially satisfy this region.
+\item $\iota_{\var{tdtp}}$, this region was constructed so the part of the heap it governs remains unaltered.
 \item $\iota_{\var{td}}$, apart from the address with the counter, this region is also constructed so this heap segment does not change.
@@
 
+We have now shown the three conditions and thus proven Lemma~\ref{lem:tckt-disp}.
 \end{proof}

### 2016-07-06 capability-machine.tex
Changed the definition of the observation relation

@@
 % Save ticket dispenser code
-\retm      & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
+\retm:     & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
            & \lea{\refreg{r_1}}{1} & & \\
@@
 % retrieve io capability
-\reta      & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
+\reta:     & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
            & \lea{r_0}{-2} & & \\
@@
 \end{figure}
+
 In the JS paper, the ticket dispenser program dereferences the counter and returns the value.
@@
 The adversary could also choose never to return to the return address that we set up for him.
-\begin{lemma}[Ticket dispenser test]
+\begin{lemma}[Ticket dispenser test]\dominique{I would prefer more precise, symbolic statements of assumptions and less prose.}
 \label{lem:tckt-disp}
@@
 Take $\hsframe$ to be everything else.
-We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.
-According to the specification of $\malloc$, we get $\Phi' \step^* \Phi''$ where
+We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.\dominique{explicit assumption about this needed?} According to the specification of $\malloc$, we get $\Phi' \step^* \Phi''$ where
 \begin{itemize}
@@
 Now we have all the necessary conditions satisfied to be able to use the fact that $c$ is in the continuation predicate to conclude
+
 \lau{did I mess up the steps here?}
@@
 %condition 2
+
+\dominique{this duplicates the argument about $(\entry,\initb,\inite,\initb+\var{ret}) \in \stdvr(W'')$. Find a way to share this?}
 For condition 2, we need to show 
@@
 Finally, we need to show $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$.
-The domains of the heap segments are assumed to be disjoined, and we have limited the heap to only be the parts considered in the regions, so the first condition of heap satisfaction is okay.
+The domains of the heap segments are assumed to be disjoint, and we have limited the heap to only be the parts considered in the regions, so the first condition of heap satisfaction is okay.
 Let us take the regions one at a time and briefly argue that the corresponding heap segment satisfies the region:
@@
                            & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
-                           & \quad (\exists i < n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
+                           & \quad (\exists i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                            & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap')  \\
-                           & \qquad \Rightarrow \npair[n-i]{\heap'(0)} \in \stdvr(W) \land \\
-                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap' \restrictfun{\{0\}}} \in W(r)(W) )\}
+                           & \qquad \Rightarrow \exists \heap'', \heap_f \ldotp \heap' = \heap'' \uplus \heap_f \land \heapSat[heap'']{n-i}{W}
 \end{align*}
-\lau{Is the last condition sufficient? If any of the other two conditions are true, then the last becomes trivially true as well.}\\
+Removed the condition $\npair[n-i]{\heap'(0)} \in \stdvr(W)$.\\
+
 The ``good observations'' are as follows
@@
 \item The execution stops with a $\failed$.
-\item The execution halts in at most $n$ steps with a valid value in in address 0 and the value at address 0 is also valid with respect to the world.
+\item The execution halts in at most $n$ steps with a heap where the island that governs address 0 is satisfied by some heap segment.
 \end{itemize}
@@
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  For any $n \in \nats$, $W \in \Worlds$, and addresses $\var{base}$ and $\var{end}$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, \var{base})} \in \stder(W)$.
+  For any $n \in \nats$, $W \in \Worlds$, and addresses $\var{base}$, $\var{end}$, and $a$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, a)} \in \stder(W)$.
 \end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)

### 2016-07-08 capability-machine.tex
Added heap satisfaction monotonicity lemma
added mono lemmas for read/write condition, added read cond to entry cond, changed updatePcPerm to be total

@@
                                      \begin{cases}
-                                       (\perm,\start,\addrend,\addr) & \text{if $\perm\in\{ \rwx, \exec \}$} \\
                                        (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}
+                                       (\perm,\start,\addrend,\addr) & \text{otherwise} \\
+
                                      \end{cases} \\
@@
   \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
+                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \land \\
+                                          & \; \quad \readCond{n',W',\addrend,\addr}
 \end{align*}
 
+\begin{lemma}[Write condition implies read condition]
+\[
+\forall n, W, \start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
+\]
+\end{lemma}
 \[
@@
                      \forall \heapSat{n}{W'}, \heap_f \ldotp \forall \reg \in \stdrr(W') \\
-                   & \qquad \npair{(reg\update{\pcreg}{\updatePcPerm{\var{c}}}, \heap \uplus \heap_f} \in \observations(W') \}
+                   & \qquad \npair{(reg\update{\pcreg}{\updatePcPerm{\var{c}}}, \heap \uplus \heap_f)} \in \observations(W') \}
 \end{align*}
@@
 \dominique{Something I find interesting here, is that the FTLR can be interpreted to say that x permissions have no real security value: if you have an arbitrary r or rw capability that is valid, then the FTLR says that it remains valid if we change it to rx or rwx (respectively).}
+
+\begin{lemma}[Heap satisfaction downwards closure]
+\[
+  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
+\]
+\end{lemma}
+
+\begin{lemma}[Read condition downwards closure]
+\[
+\forall n,n',W,\start, \addrend \ldotp \readCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \readCond{n',W,\start,\addrend}
+\]
+\end{lemma}
+\begin{lemma}[Write condition downwards closure]
+\[
+\forall n,n',W,\start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \writeCond{n',W,\start,\addrend}
+\]
+\end{lemma}
 

### 2016-07-11 capability-machine.tex
removed readCondition from entry cond

@@
 \newcommand{\readCond}[1]{\plainfun{readCondition}(#1)}
-\newcommand{\writeCond}[1]{\plainfun{writeCondition}(#1)}
+\newcommand{\writeCond}[1]{\plainfun{readWriteCondition}(#1)}
 \newcommand{\execCond}[1]{\plainfun{executeCondition}(#1)}
@@
                                      \begin{cases}
-                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}
-                                       (\perm,\start,\addrend,\addr) & \text{otherwise} \\
-
+                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}\\
+                                       (\perm,\start,\addrend,\addr) & \text{otherwise} 
                                      \end{cases} \\
@@
                                                               \text{if $r_1 \neq \pcreg$}\\
-                                                              \quad\text{and $r_1\neq \pcreg$}\\
-                                                              \quad\text{for $i \in \{1,2\}$}\\
+                                                              \quad\text{and for $i \in \{1,2\}$}\\
                                                               \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}\\
@@
                                           & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \land \\
-                                          & \; \quad \readCond{n',W',\addrend,\addr}
 \end{align*}

### 2016-07-11 capability-machine.tex
simplified semantics

@@
   \sem{\halt}(\Phi)                        & = (\halted,\Phi) \\
-  \sem{\jmp{\lv}}(\Phi)                    & = 
-                                             \begin{cases}
-                                               \updateReg{\pcreg}{\updatePcPerm{\var{c}}} &
-                                               \arraycolsep=0pt
-                                               \begin{array}{l}
-                                                 \text{if }\memreg(lv) = c \\
-                                                 \quad\text{and }c=(\perm,\start,\addrend,\addr)\\
-                                                 \quad\text{and }\executeAllowed{\perm}
-                                               \end{array} \\
-                                               \failed                              & \text{otherwise }
-                                             \end{cases} \\
+  \sem{\jmp{\lv}}(\Phi)                    & = \updateReg{\pcreg}{\updatePcPerm{\var{c}}} \\
   \sem{\jnz{\lv}{\rv}}(\Phi)               & = 
@@
                                                \begin{array}{l}
-                                                 \text{if $\nonZero{\memreg(\rv)}$} \\ 
-                                                 \quad\text{and $\memreg(lv) = c$} \\
-                                                 \quad\text{and $c=(\perm,\start,\addrend,\addr)$}\\
-                                                 \quad\text{and $\executeAllowed{\perm}$}
-                                               \end{array}
-                                               \\ %TODO Maybe combine with jump. (failed + this)
+                                                 \text{if $\nonZero{\memreg(\rv)}$} 
+                                               \end{array}\\
                                                \stdUpdatePc{\Phi} & \text{if not $\nonZero{\memreg(\rv)}$}\\
@@
                                                \begin{array}{l}
-                                                 \text{if } r_1 \neq \pcreg \\
-                                                 \quad\text{and }\memreg(r_2) = (\perm,\start,\addrend,\addr) = \var{c} \\
+                                                 \text{if }\memreg(r_2) = (\perm,\start,\addrend,\addr) = \var{c} \\
                                                  \quad\text{and }\readAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
@@
  \sem{\move{\refreg{r_1}}{\rv}}            & = \begin{cases}
-                                                 \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}} & \text{if $r_1 \neq \pcreg$} \\ % For now we do this, so you need to use a jump to move stuff into pc.
+                                                 \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}} & \\ % For now we do this, so you need to use a jump to move stuff into pc.
                                                  \failed   & \text{otherwise }
                                  \end{cases}
-\end{align*}
-\begin{align*}
+\\
   \sem{\lea{\refreg{r_1}}{\rv}}            & =
@@
                                              \end{cases} 
-\\ % In the M-Machine, lea checks whether the pointer stay within the allowed range in the pointer.
+\\
+ % In the M-Machine, lea checks whether the pointer stay within the allowed range in the pointer.
   \sem{\restrict{\refreg{r_1}}{\rv_1}{\rv_2}}           & =
@@
                                                  \begin{array}{l}
-                                                   \text{if $r_1 \neq \pcreg$}\\
-                                                   \quad\text{and $\memreg(\rv_1) = c'$}\\
+                                                   \text{if $\memreg(\rv_1) = c'$}\\
                                                    \quad\text{and $c' = \stdcap$}\\
@@
                                                \failed                   & \text{otherwise}
-                                             \end{cases} \\ 
+                                             \end{cases} 
+\end{align*}
+\begin{align*}
   \sem{\plus{\refreg{r_1}}{\rv_1}{\rv_2}}               & =
@@
                                                             \begin{array}{l}
-                                                              \text{if $r_1 \neq \pcreg$}\\
-                                                              \quad\text{and for $i \in \{1,2\}$}\\
-                                                              \qquad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}\\
-                                                              \qquad\text{and in either case $n_i \in \ints$}
+                                                              \text{if for $i \in \{1,2\}$}\\
+                                                              \quad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}\\
+                                                              \quad\text{and in either case $n_i \in \ints$}
                                                             \end{array}\\

### 2016-07-12 capability-machine.tex
added a new symbolic spec. for malloc

@@
 \lau{maybe mention that malloc is also on the heap}
- \item[$\boldsymbol \memheap$] contains the malloc code.
- \item[Nothing else is allocated in $\boldsymbol \memheap$] and all the above programs are disjoint.
+ \item[$\boldsymbol \memheap$] contains the malloc code, and the malloc code satisfy $\iota_{\malloc}$, i.e., $\forall n \ldotp \heapSat[{\heap\restrictfun{\{\malloc_{\var{base}}, \dots ,\malloc_{\var{end}}\}}}]{n}{W_{\malloc}}$.
+ \item[All the above programs are disjoint.]
  \item[$\boldsymbol{\adv}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
- \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\initb,\inite,\initb)$, which is the only capability that grants any access to this part of the heap,
- \item[$\boldsymbol{ \memreg(\rio)}$] is $(\readwrite,0,0,0)$, which also is the only capability that grants access to this part of the heap, 
+ \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\initb,\inite,\initb)$.
+ \item[$\boldsymbol{ \memreg(\rio)}$] is $(\readwrite,0,0,0)$.
  \item[the remaining registers] all contains 0,
@@
 If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
+\begin{specification}[Malloc v.2]
+  \begin{align*}
+    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments\ldotp\\
+    &\quad \forall n, i, \size \in \nats\ldotp\forall b,e,a,m_b,m_e\in \Addrs\ldotp\forall p \in \Perms \ldotp\\
+    &\qquad W_{\malloc} = [i \mapsto \iota_{\malloc}] \land \\
+    &\qquad \dom(\hsfoot) = [m_b,m_e] \land \\
+    &\qquad \memheap = \hsfoot \uplus \hsframe \land\\
+    &\qquad \heapSat[\hsfoot]{n}{W_{\malloc}} \land \\
+    &\qquad \memreg(r_1) = \size \land \\
+    &\qquad \size \geq 0 \land \\
+    &\qquad \memreg(r_0) = (p,b,e,a) \land \\
+    &\qquad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
+    &\qquad\Rightarrow \\
+    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot', \hsframe' \in \HeapSegments\ldotp\\
+    &\qquad \qquad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \\
+    &\qquad \qquad \quad \Phi \step[j]\Phi' \land \\
+    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe' \land\\
+    &\qquad \qquad \quad W' = [i \mapsto \iota_{\malloc}, k \mapsto \iota^0_{b',e'}] \land \\
+    &\qquad \qquad \quad \heapSat[\hsfoot']{n-j}{W'} \land \\
+    &\qquad \qquad \quad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
+    &\qquad \qquad \quad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
+    &\qquad \qquad \quad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
+\end{align*}
+\end{specification}
+\lau{I am not satisfied with the pc in the premise.}
+
 \begin{specification}[Malloc]\dominique{can you state this more systematically, with less prose, more explicit assumptions and conclusion?}
@@
 To use the FTLR, we have to argue that the read condition is satisfied for the region, the adversary is in.
-This is easily shown, as we have chosen the region for the adversary to be the standard one, so with $\advb$ and $\adve$ as witnesses, we need to show $\iota{\advb,\adve} \subseteq \iota{\advb,\adve}$.
+This is easily shown, as we have chosen the region for the adversary to be the standard one, so with $\advb$ and $\adve$ as witnesses, we need to show $\iota_{\advb,\adve} \subseteq \iota_{\advb,\adve}$.
 

### 2016-07-12 capability-machine.tex
removed two conditions from obs. rel.

@@
 \end{specification}
-\lau{I am not satisfied with the pc in the premise.}
+\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
 
@@
 
-If we start running in $\Phi$ until $\memreg(\pcreg)$ points to $\jmp{\malloc}$, then the configuration is the same as $\Phi$ except:
+If we start running in $\Phi$ until $\jmp{\malloc}$ has just been executed, then the configuration is the same as $\Phi$ except:
 \begin{itemize}
@@
   \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
-                           & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
-                           & \quad (\exists i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                            & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap')  \\

### 2016-07-12 capability-machine.tex
moved the framing part to the obs. relation, removed the AsmType from the relations

@@
 %LR
-\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
+\newcommand{\intr}[2]{\mathcal{#1}}
 \newcommand{\valueintr}[1]{\intr{V}{#1}}
@@
   \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
-  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
-                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap')  \\
-                           & \qquad \Rightarrow \exists \heap'', \heap_f \ldotp \heap' = \heap'' \uplus \heap_f \land \heapSat[heap'']{n-i}{W}
+  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
+                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                           & \qquad \Rightarrow \exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W}
 \end{align*}
@@
   \stdkr(W) \defeq & \{ \npair{c} \mid c \in \stdvr(W) \land \\
-                   & \quad  \forall W' \future W\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat{n}{W'}, \heap_f \ldotp \forall \reg \in \stdrr(W') \\
-                   & \qquad \npair{(reg\update{\pcreg}{\updatePcPerm{\var{c}}}, \heap \uplus \heap_f)} \in \observations(W') \}
+                   & \quad \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
+                     \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
+                   & \qquad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},hs)} \in \observations(W') \}
 \end{align*}
@@
   \stder(W) \defeq & \{ \npair{\pc} \mid \\
-                   & \quad \forall \reg \in \stdrr(W) \ldotp \\
+                   & \quad \forall \npair{\reg} \in \stdrr(W) \ldotp \\
                    & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\
-                   & \qquad \quad  \forall \heapSat{n}{W}, \heap_f \ldotp \\
-                   & \qquad \qquad \npair{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\heap \uplus \heap_f)} \in \observations(W) \}
+                   & \qquad \quad  \forall \heapSat[\hs]{n}{W} \ldotp \\
+                   & \qquad \qquad \npair{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
 \end{align*}
@@
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  For any $n \in \nats$, $W \in \Worlds$, and addresses $\var{base}$, $\var{end}$, and $a$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, a)} \in \stder(W)$.
+  For any $n \in \nats$, $W \in \Worlds$, $p\in \Perms$, and addresses $\var{base}$, $\var{end}$, and $a$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, a)} \in \stder(W)$.
 \end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)

### 2016-07-13 capability-machine.tex
Some remarks on symbolic malloc spec.
Another remark.

@@
 \lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
+\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
+  pc unspecified as it is not important for malloc users?}
+\dominique{Resulting frame should be equal to original.}
+\dominique{Allocated pointer should have the right size.}
+\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}
 

### 2016-07-26 capability-machine.tex
fixed issue in the semantics of jmp
fixed unquantified variable (c) in jnz. Added requirement for size and fixed frame issue (now it uses the same frame) in malloc specification. Updated type of observation rel. to reflect previous change.
removed impossible case in semantics
fixed typo in semantics
added new world definitions
more world definitions

@@
 \newcommand{\perm}{\var{perm}}
+\newcommand{\roll}{\var{roll}}
 
@@
 
-\newcommand{\RegionName}{\plaindom{RegionName}}
+\newcommand{\States}{\plaindom{State}}
+\newcommand{\RegionNames}{\plaindom{RegionName}}
+\newcommand{\Regions}{\plaindom{Region}}
 \newcommand{\Worlds}{\plaindom{World}}
-\newcommand{\StorePred}{\plaindom{StorePred}}
+\newcommand{\StorePred}{\plaindom{HeapSegPred}}
 \newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
@@
 %LR
-\newcommand{\intr}[2]{\mathcal{#1}}
+\newcommand{\intr}[2]{\mathcal{#1}\sem{#2}}
 \newcommand{\valueintr}[1]{\intr{V}{#1}}
@@
   \sem{\halt}(\Phi)                        & = (\halted,\Phi) \\
-  \sem{\jmp{\lv}}(\Phi)                    & = \updateReg{\pcreg}{\updatePcPerm{\var{c}}} \\
+  \sem{\jmp{\lv}}(\Phi)                    & = \updateReg{\pcreg}{\updatePcPerm{\memreg(\lv)}} \\
   \sem{\jnz{\lv}{\rv}}(\Phi)               & = 
                                              \begin{cases}
-                                               \updateReg{\pcreg}{\updatePcPerm{\var{c}}} &
+                                               \updateReg{\pcreg}{\updatePcPerm{\memreg(\lv)}} &
                                                \arraycolsep=0pt
@@
                                              \end{cases}\\
- \sem{\move{\refreg{r_1}}{\rv}}            & = \begin{cases}
-                                                 \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}} & \\ % For now we do this, so you need to use a jump to move stuff into pc.
-                                                 \failed   & \text{otherwise }
-                                 \end{cases}
+ \sem{\move{\refreg{r_1}}{\rv}}            & = \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}}
 \\
@@
                                                  \begin{array}{l}
-                                                   \text{if either $n = \rv$ or $\rv = \refheap{r_2}$ and $n = \memreg(r_2)$} \\
+                                                   \text{if either $n = \rv$ or $\rv = \refreg{r_2}$ and $n = \memreg(r_2)$} \\
                                                    \quad\text{and in either case $n \in \ints $} \\
@@
     &\qquad\Rightarrow \\
-    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot', \hsframe' \in \HeapSegments\ldotp\\
+    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
     &\qquad \qquad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \\
     &\qquad \qquad \quad \Phi \step[j]\Phi' \land \\
-    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe' \land\\
+    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
     &\qquad \qquad \quad W' = [i \mapsto \iota_{\malloc}, k \mapsto \iota^0_{b',e'}] \land \\
@@
     &\qquad \qquad \quad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
+    &\qquad \qquad \quad \size = e'-b'
     &\qquad \qquad \quad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
@@
 \[
-  \StorePred = \Worlds \monnefun \UPred{\Addrs \parfun \Words}
+  \Worlds = \{ W \in RegionNames \parfun \Regions \mid \dom(W) \text{ finite } \}
 \]
+$\States$ is a fixed set with the necessary states.
+l
+\begin{align*}
+  \Regions = \{ r = (s, \phi, H) \mid &  s \in \States \land \phi \subseteq States^2 \land \\ 
+                                      &  H : \States \rightarrow \StorePred \land \\ 
+                                      &  \phi \text{ reflexive and transitive} \}
+\end{align*}
 \[
-\Worlds \cong \frac{1}{2} (\RegionName \parfun \StorePred)
+  \StorePred = \widehat{\Worlds} \monnefun \UPred{\HeapSegments}
 \]
+
+\[
+\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
+\]
+Future Worlds:
+\[
+ W' \future W \text{ iff }  
+ \begin{gathered}
+   \dom(W') \supseteq \dom(W) \land \\
+   \forall r \in dom(W) \ldotp W'(r) \future W(r)
+ \end{gathered}
+\]
+Future regions:
+\[
+  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
+    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2 \in \phi_1)
+\]
+Heap segment satisfaction:
+\[
+\heapSat[\hs]{n}{W} \text{ iff }
+\begin{aligned}
+  \exists R \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
+  \text{and}
+  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
+\end{aligned}
+\]
+Where $\iota.s$ and $\iota.H$ project out the state and the predicate of the region respectively.
 
@@
 \[
-  \iota_{\var{start},\var{end}} \; : \; \StorePred
+  \iota_{\var{start},\var{end}} \; : \; \Regions
 \]
 \begin{align*}
-  \iota_{\var{start},\var{end}}(W) \defeq  \{ \npair{\heapseg} \mid & \dom(\heapseg) = [\var{start},\var{end}] \land \\
-                                                               & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
+  \iota_{\var{start},\var{end}}(W) & \defeq (s,=,H_{\var{std}}) \\
+  H_{\var{std}}\; s\; W & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(a) \in \stdvr(\roll \; W)\}
 \end{align*}
@@
 \begin{align*}
-  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionName \ldotp \\
+  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionNames \ldotp \\
                                           & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
-  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionName \ldotp \\
+  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionNames \ldotp \\
                                           & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
@@
 \begin{align*}
-  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
+  \observations \quad : & \quad  \Worlds \monfun \UPred{\Regs \times \HeapSegments} \\
   \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\

### 2016-07-27 capability-machine.tex
Updated malloc specification

@@
   \begin{align*}
-    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments\ldotp\\
-    &\quad \forall n, i, \size \in \nats\ldotp\forall b,e,a,m_b,m_e\in \Addrs\ldotp\forall p \in \Perms \ldotp\\
-    &\qquad W_{\malloc} = [i \mapsto \iota_{\malloc}] \land \\
-    &\qquad \dom(\hsfoot) = [m_b,m_e] \land \\
-    &\qquad \memheap = \hsfoot \uplus \hsframe \land\\
-    &\qquad \heapSat[\hsfoot]{n}{W_{\malloc}} \land \\
-    &\qquad \memreg(r_1) = \size \land \\
-    &\qquad \size \geq 0 \land \\
-    &\qquad \memreg(r_0) = (p,b,e,a) \land \\
-    &\qquad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
-    &\qquad\Rightarrow \\
-    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
-    &\qquad \qquad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \\
-    &\qquad \qquad \quad \Phi \step[j]\Phi' \land \\
-    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
-    &\qquad \qquad \quad W' = [i \mapsto \iota_{\malloc}, k \mapsto \iota^0_{b',e'}] \land \\
-    &\qquad \qquad \quad \heapSat[\hsfoot']{n-j}{W'} \land \\
-    &\qquad \qquad \quad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
-    &\qquad \qquad \quad \size = e'-b'
-    &\qquad \qquad \quad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
-    &\qquad \qquad \quad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
+    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
+    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
+    &\qquad \exists i, \var{FS} \ldotp \\
+    &\qquad \quad W (i) = \iota_{\malloc}^{\var{FS}} \land \\
+    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\
+    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
+    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
+    &\qquad \quad \memreg(r_1) = \size \land \\
+    &\qquad \quad \size \geq 0 \land \\
+    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
+    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
+    &\qquad \quad \Rightarrow \\
+    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
+    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
+    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
+    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
+    &\qquad \qquad \qquad W' \future W \land \\
+    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\var{FS} \setminus [b',e']} \land \\
+    &\qquad \qquad \qquad [b',e'] \subseteq \var{FS}  \land \\
+    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
+    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
+    &\qquad \qquad \qquad \size = e'-b' \\
+    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
+    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
 \end{align*}
@@
 \]
-Heap segment satisfaction:
+\begin{definition}[Heap satisfaction/erasure]
 \[
 \heapSat[\hs]{n}{W} \text{ iff }
-\begin{aligned}
+\begin{gathered}
   \exists R \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
-  \text{and}
+  \text{and}\\
   \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
-\end{aligned}
+\end{gathered}
 \]
 Where $\iota.s$ and $\iota.H$ project out the state and the predicate of the region respectively.
+\end{definition}
 
@@
 \begin{align*}
-  \iota_{\var{start},\var{end}}(W) & \defeq (s,=,H_{\var{std}}) \\
-  H_{\var{std}}\; s\; W & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(a) \in \stdvr(\roll \; W)\}
+  \iota_{\var{base},\var{end}}(W) & \defeq (s,=,H_{\var{base},\var{end}}) \\
+  H_{\var{base},\var{end}}\; s\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
+                             & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
 \end{align*}
@@
 \[
-  \iota_{\var{start},\var{end}}^0 \; : \; \StorePred
-\]
-\begin{align*}
-  \iota_{\var{start},\var{end}}^0(W) \defeq \{ \npair{\heapseg} \mid & \dom(\heapseg) = [\var{start},\var{end}] \land \\
+  \iota_{\var{start},\var{end}}^0 \; : \; \Regions
+\]
+\begin{align*}
+  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq (s,=,H_{\var{base},\var{end}}^0) \\
+  H_{\var{base},\var{end}}^0\; s\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                 & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
@@
 
-\begin{definition}[Heap satisfaction/erasure]
-\begin{align*}
-  \heapSat{n}{W} & & \text{iff} & &
-                                      \arraycolsep=0pt
-                                      \begin{array}{l}
-                                      \exists r: \dom(W) \rightarrow \HeapSegments \ldotp \\
-                                      \qquad\heap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
-                                      \quad\text{and}\\
-                                      \qquad\forall \iota \in \dom(W) \ldotp \npair{r(\iota)} \in W(\iota)(W)
-                                      \end{array}\\
-\end{align*}
-
-\end{definition}
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline

### 2016-08-01 capability-machine.tex
Added well-definedness lemmas for the logical predicates.

@@
 \newcommand{\reta}{\var{ret}_{\adv}}
+\newcommand{\FS}{\var{FS}}
 \newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
@@
     &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
-    &\qquad \exists i, \var{FS} \ldotp \\
-    &\qquad \quad W (i) = \iota_{\malloc}^{\var{FS}} \land \\
+    &\qquad \exists i, \FS \ldotp \\
+    &\qquad \quad W (i) = \iota_{\malloc}^{\FS} \land \\
     &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\
@@
     &\qquad \qquad \qquad W' \future W \land \\
-    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\var{FS} \setminus [b',e']} \land \\
-    &\qquad \qquad \qquad [b',e'] \subseteq \var{FS}  \land \\
+    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\FS \setminus [b',e']} \land \\
+    &\qquad \qquad \qquad [b',e'] \subseteq \FS  \land \\
     &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
@@
 \end{specification}
+The specification uses the mallog island defined as
+\[
+\begin{gathered}
+  \iota_{\malloc}^{\FS} \defeq (\FS, \phi_{\malloc}, H_{\malloc}) \\
+  \FS \subseteq \Addrs \\
+  \forall \FS_1, \FS_2 \ldotp (\FS_1,\FS_2) \in \phi_{\malloc} \text{ iff } \FS_1 \supseteq \FS_2 \\
+  H \; W \; \FS = {\npair{\hs} \mid \dom(\hs) = \FS}
+\end{gathered}
+\]
 \lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
@@
 $\States$ is a fixed set with the necessary states.
-l
+
 \begin{align*}
@@
 \end{align*}
+For convenience, we write $r.s$, $s.\phi$, and $s.H$ for a the projection of a regions state, transition system, and store predicate, respectively
 \[
@@
   (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
-    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2 \in \phi_1)
+    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2) \in \phi_1
 \]
@@
 \heapSat[\hs]{n}{W} \text{ iff }
+\exists R \ldotp 
 \begin{gathered}
-  \exists R \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
+  \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
   \text{and}\\
@@
 \]
-Where $\iota.s$ and $\iota.H$ project out the state and the predicate of the region respectively.
 \end{definition}
+
+\begin{lemma}[Heap satisfaction uniformity]
+\[
+  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
+\]
+\end{lemma}
 
@@
 
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
+  \]
+\end{lemma}
+
 Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
@@
                            & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
-                           & \qquad \Rightarrow \exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W}
+                           & \qquad \Rightarrow \exists W' \future W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
 \end{align*}
@@
 Address 0 is the designated I/O address.
+\\ \lau{The two first conditions where implied by the last one, so they were removed.
+It turned out to cause troubles to special case on address 0, so now the entire heap has to satisfy a future world (which is a bit worrying as only the i/o address is observable).}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W \future W' \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair{(\reg,\hs)}(W')
+  \]
+\end{lemma}
 
@@
 \end{align*}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
+  \]
+\end{lemma}
 
@@
 Here everything that is in the register file is available when the continuation is invoked, so we do not special case on $r_1$ (which is the register that in the calling convention we use should contain the return value).
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
+  \]
+\end{lemma}
 
@@
 \end{align*}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall \pc \ldotp \npair{c} \in \stder(W) \Rightarrow \npair{c} \in \stder(W')
+  \]
+\end{lemma}
 
@@
 
-\begin{lemma}[Heap satisfaction downwards closure]
-\[
-  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
-\]
-\end{lemma}
-
-\begin{lemma}[Read condition downwards closure]
+\begin{lemma}[Read condition uniformity]
 \[
@@
 \end{lemma}
-\begin{lemma}[Write condition downwards closure]
+
+\begin{lemma}[Write condition uniformity]
 \[

### 2016-08-02 capability-machine.tex
Placed well formedness lemmas in the end of the document
Updated the malloc specification.

@@
 %Comments
-\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}}
-\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}}
+\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
+\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
 
@@
 %Functions
-\newcommand{\plainfun}[1]{\mathit{#1}}
-\newcommand{\decode}{\plainfun{decode}}
-\newcommand{\encode}{\plainfun{encode}}
-\newcommand{\encodePerm}{\plainfun{encodePerm}}
-\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}(#1)}
-
-\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}(#1)}
-\newcommand{\nonZero}[1]{\plainfun{nonZero}(#1)}
-\newcommand{\readAllowed}[1]{\plainfun{readAllowed}(#1)}
-\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}(#1)}
-\newcommand{\withinBounds}[1]{\plainfun{withinBounds}(#1)}
-\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc(#1)}}
-
-\newcommand{\readCond}[1]{\plainfun{readCondition}(#1)}
-\newcommand{\writeCond}[1]{\plainfun{readWriteCondition}(#1)}
-\newcommand{\execCond}[1]{\plainfun{executeCondition}(#1)}
-\newcommand{\entryCond}[1]{\plainfun{entryCondition}(#1)}
+\newcommand{\plainfun}[2]{
+  \ifthenelse{\equal{#2}{}}
+             {\mathit{#1}}
+             {\mathit{#1}(#2)}
+}
+\newcommand{\decode}{\plainfun{decode}{}}
+\newcommand{\encode}{\plainfun{encode}{}}
+\newcommand{\encodePerm}{\plainfun{encodePerm}{}}
+\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}
+
+\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}{#1}}
+\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
+\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
+\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
+\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
+\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc}{#1}}
+
+\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
+\newcommand{\writeCond}[1]{\plainfun{readWriteCondition}{#1}}
+\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
+\newcommand{\entryCond}[1]{\plainfun{entryCondition}{#1}}
 
@@
     &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
-    &\qquad \exists i, \FS \ldotp \\
-    &\qquad \quad W (i) = \iota_{\malloc}^{\FS} \land \\
-    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\
+    &\qquad \exists i, \iota_{\malloc} \ldotp \\
+    &\qquad \quad W = [i \mapsto \iota_{\malloc}] \land \\
+    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\ % This
     &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
@@
     &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
-    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
+    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\ %and this needs to change with ABI.
     &\qquad \quad \Rightarrow \\
@@
     &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
+    &\qquad \qquad \qquad W' = [k \mapsto \iota_{b',e'}^0, i \mapsto \iota_{\malloc}]\\
     &\qquad \qquad \qquad W' \future W \land \\
-    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\FS \setminus [b',e']} \land \\
     &\qquad \qquad \qquad [b',e'] \subseteq \FS  \land \\
@@
 The specification uses the mallog island defined as
-\[
-\begin{gathered}
-  \iota_{\malloc}^{\FS} \defeq (\FS, \phi_{\malloc}, H_{\malloc}) \\
-  \FS \subseteq \Addrs \\
-  \forall \FS_1, \FS_2 \ldotp (\FS_1,\FS_2) \in \phi_{\malloc} \text{ iff } \FS_1 \supseteq \FS_2 \\
-  H \; W \; \FS = {\npair{\hs} \mid \dom(\hs) = \FS}
-\end{gathered}
-\]
 \lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
@@
 \dominique{Resulting frame should be equal to original.}
-\dominique{Allocated pointer should have the right size.}
-\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}
+\lau{Rewrite the specification when we have settled on a reasonable way to handle ABI.}
 
@@
 
-\begin{lemma}[Heap satisfaction uniformity]
-\[
-  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
-\]
-\end{lemma}
-
 A standard heap invariant that insures all values in the region are in the value  relation:
@@
 
-\begin{lemma}[Write condition implies read condition]
-\[
-\forall n, W, \start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
-\]
-\end{lemma}
 \[
@@
 
-\begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
-  \]
-\end{lemma}
-
 Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
@@
 
-\begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n \ldotp \forall W \future W' \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair{(\reg,\hs)}(W')
-  \]
-\end{lemma}
-
 Register-file relation:
@@
 \end{align*}
-
-\begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
-  \]
-\end{lemma}
 
@@
 Here everything that is in the register file is available when the continuation is invoked, so we do not special case on $r_1$ (which is the register that in the calling convention we use should contain the return value).
-
-\begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
-  \]
-\end{lemma}
 
@@
 \end{align*}
-
-\begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W)
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall \pc \ldotp \npair{c} \in \stder(W) \Rightarrow \npair{c} \in \stder(W')
-  \]
-\end{lemma}
 
@@
 
+\begin{lemma}[Heap satisfaction uniformity]
+\[
+  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
+\]
+\end{lemma}
+
 \begin{lemma}[Read condition uniformity]
@@
 \end{lemma}
-
-\begin{lemma}[Write condition uniformity]
+\begin{proof}
+
+\end{proof}
+
+\begin{lemma}[Read-write condition uniformity]
 \[
 \forall n,n',W,\start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \writeCond{n',W,\start,\addrend}
+\]
+\end{lemma}
+
+\begin{lemma}[Execute condition uniformity]
+\[
+\forall n,n',W,\start, \addrend, \perm \ldotp \execCond{n,W,\start,\addrend,\perm} \land n' \leq n \Rightarrow \execCond{n',W,\start,\addrend,\perm}
+\]
+\end{lemma}
+
+\begin{lemma}[Entry condition uniformity]
+\[
+\forall n,n',W,\start, \addrend, \addr \ldotp \entryCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \entryCond{n',W,\start,\addrend,\addr}
+\]
+\end{lemma}
+
+\begin{lemma}[Write condition implies read condition]
+\[
+\forall n, W, \start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
+\]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
+  \]
+\end{lemma}
+\begin{proof}
+  Follows from the uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$.
+\end{proof}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W \future W' \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair{(\reg,\hs)}(W')
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
+  \[
+    \forall n \ldotp \forall W' \future W \ldotp \forall \pc \ldotp \npair{c} \in \stder(W) \Rightarrow \npair{c} \in \stder(W')
   \]

### 2016-08-04 capability-machine.tex
Added short proofs for a couple of the lemmas.

@@
 \newcommand{\reta}{\var{ret}_{\adv}}
-\newcommand{\FS}{\var{FS}}
+\newcommand{\base}{\var{base}}
+\newcommand{\eend}{\var{end}}
 \newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
@@
     &\qquad \qquad \qquad W' \future W \land \\
-    &\qquad \qquad \qquad [b',e'] \subseteq \FS  \land \\
     &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
@@
 \begin{align*}
-  \Regions = \{ r = (s, \phi, H) \mid &  s \in \States \land \phi \subseteq States^2 \land \\ 
+  \Regions = \{ r = (s, \phi, H) \mid &  s \in \States \land \phi \subseteq \States^2 \land \\ 
                                       &  H : \States \rightarrow \StorePred \land \\ 
@@
 \heapSat[\hs]{n}{W} \text{ iff }
-\exists R \ldotp 
-\begin{gathered}
-  \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
+\begin{aligned}
+&\exists R : \dom(W) \rightarrow \HeapSegments \ldotp \\
+&\quad\begin{gathered}
+   hs = \biguplus_{r \in \dom(W)} R(r) \\
   \text{and}\\
@@
 \end{gathered}
+\end{aligned}
 \]
@@
 \begin{align*}
-  \observations \quad : & \quad  \Worlds \monfun \UPred{\Regs \times \HeapSegments} \\
+  \observations \quad : & \quad  \Worlds \rightarrow \UPred{\Regs \times \HeapSegments} \\
   \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
@@
 \begin{align*}
-  \stder \quad : & \quad \Worlds \monfun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
+  \stder \quad : & \quad \Worlds \rightarrow \UPred{\Words} \\%TODO: Check whether this needs to be monotone
   \stder(W) \defeq & \{ \npair{\pc} \mid \\
@@
 \begin{lemma}[Heap satisfaction uniformity]
-\[
-  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
-\]
+  \begin{align*}
+    & \forall n,n',\heap, W \ldotp\\
+    & \quad \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}[Read condition uniformity]
-\[
-\forall n,n',W,\start, \addrend \ldotp \readCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \readCond{n',W,\start,\addrend}
-\]
+  \begin{align*}
+    &\forall n,n',W,\start, \addrend \ldotp \\
+    &\quad  \readCond{n,W,\start,\addrend} \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow \readCond{n',W,\start,\addrend}
+  \end{align*}
 \end{lemma}
 \begin{proof}
-
+  Follows directly from definition: Let $n$, $n'$, $\start$, and  $\addrend$ be given.
+Assume $\readCond{n,W,\start,\addrend}$ and $n' \leq n$.
+If $n=n'$, then the result follows trivially, as it is exactly our assumption.
+If $n' < n$, then by assumption we have a region name $r$, interval $[\start',\addrend'] \supseteq [\start,\addrend]$, and the assumption $W(r) \nsubeq[n-1] \iota_{\start',\addrend'}$  given by the read condition assumption.
+Picking $r$ and $[\start',\addrend']$, we need to show  $W(r) \nsubeq[n'-1] \iota_{\start',\addrend'}$ which follows from downwards closure of $\nsubeq$.
 \end{proof}
@@
 \begin{lemma}[Read-write condition uniformity]
-\[
-\forall n,n',W,\start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \writeCond{n',W,\start,\addrend}
-\]
-\end{lemma}
+  \begin{align*}
+    & \forall n,n',W,\start, \addrend \ldotp\\
+    &\quad \writeCond{n,W,\start,\addrend} \land\\
+    &\quad  n' \leq n\\
+    &\qquad \Rightarrow \writeCond{n',W,\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Follows directly from the definition.
+\end{proof}
 
 \begin{lemma}[Execute condition uniformity]
-\[
-\forall n,n',W,\start, \addrend, \perm \ldotp \execCond{n,W,\start,\addrend,\perm} \land n' \leq n \Rightarrow \execCond{n',W,\start,\addrend,\perm}
-\]
-\end{lemma}
+  \begin{align*}
+    &\forall n,n',W,\start, \addrend, \perm \ldotp \\
+    &\quad  \execCond{n,W,\start,\addrend,\perm} \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow \execCond{n',W,\start,\addrend,\perm}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Follows directly from definition: Let $n$, $n'$, $W$, $\start$, and $\addrend$ be given.
+Assume $\execCond{n,W,\start,\addrend,\perm}$ and $n' \leq n$.
+If $n' = n$, then the result follows directly from assumption.
+If $n' < n$, then let $n'' < n'$, $W' \future W$, and $a \in [\start,\addrend]$ be given.
+Using that we by transitivity have $n'' < n$, we get the desired result from assumption.
+\end{proof}
 
 \begin{lemma}[Entry condition uniformity]
-\[
-\forall n,n',W,\start, \addrend, \addr \ldotp \entryCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \entryCond{n',W,\start,\addrend,\addr}
-\]
-\end{lemma}
+  \begin{align*}
+    & \forall n,n',W,\start, \addrend, \addr \ldotp\\
+    &\quad  \entryCond{n,W,\start,\addrend,\addr} \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow \entryCond{n',W,\start,\addrend,\addr}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Like the $\execCond{}$ uniformity proof.
+\end{proof}
 
 \begin{lemma}[Write condition implies read condition]
-\[
-\forall n, W, \start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
-\]
-\end{lemma}
+  \begin{align*}
+    &\forall n, W, \start, \addrend \ldotp\\ 
+    &\quad \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Follows directly from the definition.
+\end{proof}
 
 \begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
-  \]
+  \begin{align*}
+    &\forall n' < n \ldotp \forall W \ldotp \forall w \ldotp\\
+    &\quad \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
-  \]
+  \begin{align*}
+    &\forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp\\ 
+    &\quad \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
-  \]
+  \begin{align*}
+    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp\\
+    &\quad \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n \ldotp \forall W \future W' \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair{(\reg,\hs)}(W')
-  \]
+  \begin{align*}
+    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp\\
+    &\quad \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
-  \]
+  \begin{align*}
+    &\forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp\\
+    &\quad \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
-  \]
+  \begin{align*}
+    &\forall n' < n \ldotp \forall W \ldotp \forall c \ldotp\\
+    &\quad \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
-  \]
+  \begin{align*}
+    &\forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp\\
+    &\quad \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
+  \end{align*}
 \end{lemma}
@@
 \begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W)
-  \]
-\end{lemma}
-
-\begin{lemma}
-  \[
-    \forall n \ldotp \forall W' \future W \ldotp \forall \pc \ldotp \npair{c} \in \stder(W) \Rightarrow \npair{c} \in \stder(W')
-  \]
+  \begin{align*}
+    & \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp\\
+    & \quad \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W)
+  \end{align*}
 \end{lemma}

### 2016-08-05 capability-machine.tex
.

@@
 \]
+
+n-subset for regions
+\[
+(s_1,\phi_1,H_1) \nsubeq (s_2,\phi_2,H_2) \text{ iff } 
+\begin{gathered}
+(s_1,\phi_1) = (s_2,\phi_2)\\
+ \text{ and }\\
+ \forall \hat{W} \ldotp H_1 \; s_1 \; \hat{W} \nsubeq H_2 \; s_2 \; \hat{W}
+\end{gathered}
+\]
 \begin{definition}[Heap satisfaction/erasure]
@@
 \begin{align*}
-  \iota_{\var{base},\var{end}}(W) & \defeq (s,=,H_{\var{base},\var{end}}) \\
-  H_{\var{base},\var{end}}\; s\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
+  \iota_{\var{base},\var{end}}(W) & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
+  H_{\var{std}}\; (\var{base},\var{end})\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
                              & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
 \end{align*}
+where $s$ is some state. It does not matter which one because the predicate does not use. 
+% This is unlike the \lambda_JS paper where the state in the standard island is a location, namely the location that the standard island governs.
+
 Another standard heap invariant that ensures all the words in the region are 0.
@@
 \begin{align*}
-  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq (s,=,H_{\var{base},\var{end}}^0) \\
-  H_{\var{base},\var{end}}^0\; s\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
+  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq ((\var{base},\var{end}),=,H_{\var{std}}^0) \\
+  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                 & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
@@
   \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionNames \ldotp \\
-                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
+                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp\\
+                                          & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'}(r) \\ & \\
   \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionNames \ldotp \\
@@
 
+\subsection{Well-foundedness lemmas}
 \begin{lemma}[Heap satisfaction uniformity]
@@
 
+\begin{lemma}[Read condition monotone in world]
+  \begin{align*}
+    &\forall n,W,W',\start, \addrend \ldotp \\
+    &\quad  \readCond{n,W,\start,\addrend} \land \\
+    &\quad  W' \future W \\
+    &\qquad \Rightarrow \readCond{n,W',\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+Let $n$,$W$, $W'$, $b$, and $e$ be given where $W' \future W$.
+Assume $\readCond{n,W,\start,\addrend}$, this give a $r$ and $[b',e'] \supseteq [b,e]$ such that $W(r) \nsubeq \iota_{b',e'} (r) = ((b',e'),=,H_{\var{std}})$.
+This gives
+\begin{align*}
+  &W(r).\phi = = \\
+  &W(r).s    = (b',e') \\
+  &\forall \hat{W} \ldotp W(r).H\; (b',e') \; \hat{W} \nsubeq[n-1] H_{\var{std}}\; (b',e') \; \hat{W}
+\end{align*}
+$W' \future W$ gives
+\begin{align*}
+  &W'(r).\phi = W(r).phi \\
+  &(W'(r).s,W(r).s) \in W(r).\phi \\
+  &W'(r).H = W(r).H \\
+ \end{align*}
+Combining these two sets of properties line for line, we get the following for $W'$:
+\begin{align*}
+  & W'(r).\phi = = \\
+  & W'(r) = (b',e') \\
+  &\forall \hat{W} \ldotp W'(r).H\; (b',e') \; \hat{W} \nsubeq[n-1] H_{\var{std}}\; (b',e') \; \hat{W}
+\end{align*}
+These are exactly the properties neede to get $W'(r) \nsubeq \iota_{b',e'} (r)$ which is exactly the result we want.
+\end{proof}
+
 \begin{lemma}[Read-write condition uniformity]
@@
   Follows directly from the definition.
+\end{proof}
+
+\begin{lemma}[Read-write condition monotone in world]
+  \begin{align*}
+    & \forall n,W, W',\start, \addrend \ldotp\\
+    &\quad \writeCond{n,W,\start,\addrend} \land\\
+    &\quad  W' \leq W\\
+    &\qquad \Rightarrow \writeCond{n,W',\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Similar to the proof of the monotonicity of the read condition.
 \end{proof}
@@
 
+\begin{lemma}[Execute condition uniformity]
+  \begin{align*}
+    &\forall n,W, W',\start, \addrend, \perm \ldotp \\
+    &\quad  \execCond{n,W,\start,\addrend,\perm} \land \\
+    &\quad  W' \leq W \\
+    &\qquad \Rightarrow \execCond{n,W',\start,\addrend,\perm}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Follows directly from definition.
+\end{proof}
+
 \begin{lemma}[Entry condition uniformity]
@@
   Like the $\execCond{}$ uniformity proof.
+\end{proof}
+
+\begin{lemma}[Entry condition uniformity]
+  \begin{align*}
+    & \forall n,W,W',\start, \addrend, \addr \ldotp\\
+    &\quad  \entryCond{n,W,\start,\addrend,\addr} \land \\
+    &\quad  W' \leq W \\
+    &\qquad \Rightarrow \entryCond{n,W',\start,\addrend,\addr}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Follows directly from definition
 \end{proof}

### 2016-08-15 capability-machine.tex
changed the expression relation to be forall smaller n'

@@
   \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \land \\
+                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
 \end{align*}
@@
   \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: Check if this is still monotone
-  \stdkr(W) \defeq & \{ \npair{c} \mid c \in \stdvr(W) \land \\
+  \stdkr(W) \defeq & \{ \npair{c} \mid \npair{c} \in \stdvr(W) \land \\
                    & \quad \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg \in \stdrr(W') \\
+                     \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(W') \ldotp \\
                    & \qquad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},hs)} \in \observations(W') \}
@@
   \stder(W) \defeq & \{ \npair{\pc} \mid \\
-                   & \quad \forall \npair{\reg} \in \stdrr(W) \ldotp \\
-                   & \qquad \forall \npair{c} \in \stdkr(W) \ldotp \\
-                   & \qquad \quad  \forall \heapSat[\hs]{n}{W} \ldotp \\
-                   & \qquad \qquad \npair{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
-\end{align*}
+                   & \quad \forall n' \leq n\\
+                   & \qquad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
+                   & \qquad \quad \forall \npair[n']{c} \in \stdkr(W) \ldotp \\
+                   & \qquad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
+                   & \qquad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
+\end{align*}
+\lau{Added $n' \leq n$ to be able to prove Lemma~\ref{lem:stder-mon-step}}
 
@@
 \begin{lemma}
+\label{lem:stder-mon-step}
   \begin{align*}

### 2016-08-16 capability-machine.tex
typos and added proofs for some of the well-foundedness lemmas.

@@
 \end{align*}
-For convenience, we write $r.s$, $s.\phi$, and $s.H$ for a the projection of a regions state, transition system, and store predicate, respectively
+For convenience, we write $r.s$, $s.\phi$, and $s.H$ for a the projection of a regions state, transition system, and store predicate, respectively.
 \[
@@
  \begin{gathered}
-   \dom(W') \supseteq \dom(W) \land \\
+   \dom(W') \supseteq \dom(W)\\ 
+   \text{ and } \\
    \forall r \in dom(W) \ldotp W'(r) \future W(r)
@@
 \]
-
-n-subset for regions
+n-subset for regions:
 \[
@@
 \begin{align*}
-  \iota_{\var{base},\var{end}}(W) & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
+  \iota_{\var{base},\var{end}} & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
   H_{\var{std}}\; (\var{base},\var{end})\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
@@
 \end{align*}
-where $s$ is some state. It does not matter which one because the predicate does not use. 
-% This is unlike the \lambda_JS paper where the state in the standard island is a location, namely the location that the standard island governs.
 
@@
 \begin{align*}
-  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq ((\var{base},\var{end}),=,H_{\var{std}}^0) \\
-  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
+  \iota_{\var{base},\var{end}}^0 & \defeq ((\var{base},\var{end}),=,H_{\var{std}}^0) \\
+  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{base},\var{end}] \land \\
                                                                 & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
@@
 
-Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
 \begin{align*}
@@
   \stder(W) \defeq & \{ \npair{\pc} \mid \\
-                   & \quad \forall n' \leq n\\
+                   & \quad \forall n' \leq n \ldotp\\
                    & \qquad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
@@
 \subsection{Well-foundedness lemmas}
+\begin{lemma}
+  \begin{align*}
+    & \forall W, W', W'' \ldotp W'' \future W' \land W' \future W \Rightarrow W'' \future W
+  \end{align*}
+\end{lemma}
+
 \begin{lemma}[Heap satisfaction uniformity]
@@
 \begin{lemma}[Read condition monotone in world]
+\label{lem:readCond-mono-world}
   \begin{align*}
@@
 \begin{lemma}[Read-write condition monotone in world]
+\label{lem:writeCond-mono-world}
   \begin{align*}
@@
 
-\begin{lemma}[Execute condition uniformity]
+\begin{lemma}[Execute condition monotone in world]
+\label{lem:execCond-mono-world}
   \begin{align*}
@@
 
-\begin{lemma}[Entry condition uniformity]
+\begin{lemma}[Entry condition monotone in world]
+  \label{lem:entryCond-mono-world}
   \begin{align*}
@@
 \begin{proof}
-  Follows directly from definition
+  Follows directly from definition.
 \end{proof}
@@
 \begin{lemma}
+\label{lem:stdvr-uniform-step}
   \begin{align*}
@@
 \end{lemma}
+\begin{proof}
+  Follows from uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond$ in the worlds.
+That is Lemma~\ref{lem:readCond-mono-world},~\ref{lem:writeCond-mono-world},~\ref{lem:execCond-mono-world}, and~\ref{lem:entryCond-mono-world}.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Let $n' < n$, $W$, $\reg$, and $\hs$ be given.
+Assume $\npair{(\reg,\hs)} \in \observations(W)$.
+Let $\heap_f$, $heap'$, $i \leq n'$ be given and assume $(\reg, \hs \uplus \heap_f) \step[i] (\halted, \heap')$.
+By assumption, we have a $W' \future W$ and $hs'$ such that
+  \begin{align}
+&heap' = hs' \uplus \heap_f \label{lem:eq:stdvr1}\\
+&\heapSat[hs']{n-i}{W'}\label{lem:eq:stdvr2}
+  \end{align}. 
+Using $W'$ and $hs'$ as existential witnesses, we already have Equation~\ref{lem:eq:stdvr1} as the first necessary condition and from the above heap satisfaction along with heap satisfaction being uniform in $n$, we get $\heapSat[\hs']{n'-i}{W'}$.
+These are the two conditions necessary to get $\npair[n']{(\reg,\hs)} \in \observations(W)$.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Follows by uniformity of the value relation, Lemma~\ref{lem:stdvr-uniform-step}.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  This follows by monotonicity of $\stdvr$ in the worlds.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+Follows by using the assumption and transitivity of inequality.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+We need to show two things: The continuation is in the value predicate in the future world and that continuing execution using the continuation is in the observation predicate.
+
+The first follows from the value predicate being monotone in the world.
+
+The second property follows by assumption and transitivity of future worlds.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Follows from the assumption.
+\end{proof}
 

### 2016-08-22 capability-machine.tex
Work in progress on FTLR proof, other small corrections to definitions

@@
 \newcommand{\roll}{\var{roll}}
+\newcommand{\instr}{\var{instr}}
 
@@
 %No arguments
-\newcommand{\fail}{\instr{fail}}
-\newcommand{\halt}{\instr{halt}}
+\newcommand{\zinstr}[1]{\mathtt{#1}}
+\newcommand{\fail}{\zinstr{fail}}
+\newcommand{\halt}{\zinstr{halt}}
 %One argument
-\newcommand{\instr}[1]{\mathtt{#1}}
-\newcommand{\oneinstr}[2]{\instr{#1} \; #2}
+\newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
 \newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
 %Two arguments
-\newcommand{\twoinstr}[3]{\instr{#1} \; #2 \; #3}
+\newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
 \newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
@@
 %Three arguments
-\newcommand{\threeinstr}[4]{\instr{#1} \; #2 \; #3 \; #4}
+\newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
 \newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
@@
 \begin{align*}
-\decode &:\ints \rightarrow \Instrs
+\decode &:\Words \rightarrow \Instrs
 \end{align*}
@@
 \begin{align*}
-  \Phi & \rightarrow \sem{\decode(\memreg(\pcreg))}(\Phi) & &                                   
+  \Phi & \rightarrow \sem{\decode(\memheap(\addr))}(\Phi) & &                                   
                                                               \arraycolsep=0pt
@@
 % retrieve io capability
-\reta:     & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capabiltiy} \\
+\reta:     & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capability} \\
            & \lea{r_0}{-2} & & \\
@@
 \end{specification}
-The specification uses the mallog island defined as
-\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
+The specification uses the malloc island defined as
+\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depend on the implementation (this is mostly a note to myself.)}
 \dominique{I agree with the remark about the pc.  Perhaps leave the form of the
@@
 As part of the heap satisfies a future world of $W''$, we know that the ticket dispenser test program remains unaltered.
-If we run it (and do not run out of steps), then it loads the couner, stores it to address $0$, and halts.
+If we run it (and do not run out of steps), then it loads the counter, stores it to address $0$, and halts.
 From the heap satisfaction, we get that the counter is positive and even, so the value we store to address 0 is positive and even.
@@
 \item $\iota_{\malloc}$, from the specification of $\malloc$ we got that the resulting heap after the allocation satisfied a world with $\iota_{\malloc}$.
-We use the same region and we have not changed that part of the heap, so it still satifies this region.
+We use the same region and we have not changed that part of the heap, so it still satisfies this region.
 \end{itemize}
@@
    \text{ and } \\
-   \forall r \in dom(W) \ldotp W'(r) \future W(r)
+   \forall r \in \dom(W) \ldotp W'(r) \future W(r)
  \end{gathered}
@@
 &\quad\begin{gathered}
-   hs = \biguplus_{r \in \dom(W)} R(r) \\
+  \hs = \biguplus_{r \in \dom(W)} R(r) \\
   \text{and}\\
@@
   \iota_{\var{base},\var{end}} & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
-  H_{\var{std}}\; (\var{base},\var{end})\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
+  H_{\var{std}}\; (\var{base},\var{end})\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\roll \; W) \\
                              & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
@@
                      \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(W') \ldotp \\
-                   & \qquad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},hs)} \in \observations(W') \}
+                   & \qquad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
 \end{align*}
@@
 \dominique{Something I find interesting here, is that the FTLR can be interpreted to say that x permissions have no real security value: if you have an arbitrary r or rw capability that is valid, then the FTLR says that it remains valid if we change it to rx or rwx (respectively).}
+\begin{proof}
+  The proof is by induction on $n$.
+
+If $n=0$, then it all boils down to an assumption where $(\reg, \hs \uplus \heap_f) \step[0] (\halted, \heap')$ for some $\reg$, $\hs$, $\heap_f$, and $\heap'$ which is not possible in our operational semantics.
+
+If $n > 0$, then we need to consider a lot of cases, but first we make a lot of preliminary assumptions:\\
+Let the world $W$, permission $p$, and addresses $b$, $e$, and $a$ be given. Assume
+\begin{align}
+  & p = \rwx \text{ and } \writeCond{n,W,b,e} \text{ or } \label{eq:ftlr:rwx-assump}\\
+  & p = \readonly \text{ and } \readCond{n,W,b,e} \label{eq:ftlr:rx-assump}
+\end{align}
+We need to show 
+\[
+  \npair{(p,b,e,a)} \in \stder(W)
+\]
+Let $n'$, $\reg$, $c$, $\hs$ be given where
+\begin{align}
+  & n' \leq n \label{eq:ftlr:n-leq} \\
+  & \npair[n']{\reg} \in \stdrr(W) \label{eq:ftlr:reg-in-stdrr}\\
+  & \npair[n']{c} \in \stdkr(W) \label{eq:ftlr:cont-in-stdkr}\\
+  & \heapSat[\hs]{n'}{W} \label{eq:ftlr:hs-heapsat}
+\end{align}
+For $\pc = (p,b,e,a)$ we need to show
+\[
+\npair[n']{(\reg[r_0 \mapsto c, \pcreg \mapsto \pc],\hs)} \in \observations(W)
+\]
+For convenience, we will define $\reg' = \reg[r_0 \mapsto c, \pcreg \mapsto \pc]$. To show the above, let $h_f$, $h'$, and $i$ be given where $i \leq n'$ and assume
+\begin{equation}
+  \label{eq:ftlr:conf-steps}
+  \Phi \step[i] (\halted, h')
+\end{equation}
+for $\Phi = (\reg',\hs \uplus h_f)$ then we need to show
+\begin{align}
+  & \exists W' \future W \ldotp \exists \hs' \ldotp \nonumber \\
+  & \quad h' =  \hs' \uplus h_f \text{ and } \label{eq:ftlr:final-heap-cond} \\
+  & \quad \heapSat[\hs']{n'-i}{W'} \label{eq:ftlr:heap-sat-cond}
+\end{align}
+
+If $i=0$, then we are trivially done as execution~\ref{eq:ftlr:conf-steps} is not possible.
+If $i>0$, then execution~\ref{eq:ftlr:conf-steps} takes at least one step, consider the first step of this execution.
+By the operational semantics, the first step is:
+\begin{equation*}
+  \Phi \step \sem{\decode(h(a))}(\Phi)
+\end{equation*}
+
+where $h = \hs \uplus h_f$. We may further assume
+\begin{align}
+  &\reg'(\pcreg) = (\perm,\start,\addrend,\addr) \nonumber \\
+  &\start \leq a \leq  \addrend \nonumber \\
+  &\perm \in \{\exec,\rwx\} \nonumber
+\end{align}
+For convenience define $\instr = \decode(h(a))$. We proceed by case on $\instr$.
+\newcommand{\case}[1]{{\bf Case } $\instr=#1$:\\}
+
+\case{\fail} This case is trivially true as it is not possible to step to a halting state from a failed step in any number of steps.
+By assumption~\ref{eq:ftlr:conf-steps}, the execution ends in a halting state.
+
+\case{\halt} 
+Here we have:
+\[
+  \Phi \step (\halted, (\reg',h))
+\]
+Pick $\hs' = \hs$ and $W' = W$.
+$h' = \hs \uplus h_f$, so condition~\ref{eq:ftlr:final-heap-cond} is satisfied.
+Using $\heapSat[\hs]{n'}{W}$ and the uniformity of heap satisfaction, we get $\heapSat[\hs]{n'-i}{W'}$ as $n' > n'-i$.
+
+\case{\jmp{r}}
+Here the first execution step is:
+\[
+  \Phi \step (\reg[r_0 \mapsto c, \pcreg \mapsto \updatePcPerm{\reg'(r)}],h)
+\]
+We now consider four cases 
+\begin{description}
+\item[Case $\reg'(r) = (\rwx,b',e',a')$:] In this case, we want to appeal to the induction hypothesis.
+To do so, we need to argue that we have $\writeCond{n'-1,b',e',a'}$.
+If $r=\pcreg$, then we have this by assumption \ref{eq:ftlr:rwx-assump} and uniformity.
+If $r \neq \pcreg$, then if $r= r_0$, then by assumption \ref{eq:ftlr:cont-in-stdkr}, we have $\npair[n']{c} \in \stdkr(W)$ which entails $\npair[n']{c} \in \stdvr(W)$ which together with uniformity gives the $\writeCond{}$ due to the permission.
+Otherwise, the same folloes from assumption \ref{eq:ftlr:reg-in-stdrr} and uniformity.
+
+By IH we get 
+\begin{equation}
+  \npair[n'-1]{(\rwx,b',e',a')} \in \stder(W) \nonumber
+\end{equation}
+using $n'-1$, $\reg$, $c$, and $\hs$ and assumptions \ref{eq:ftlr:n-leq} - \ref{eq:ftlr:hs-heapsat} and the appropriate uniformity lemmas, we can use the above to get
+\begin{equation}
+  \npair[n'-1]{reg[r_0 \mapsto c, \pcreg \mapsto (\rwx,b',e',a')]} \in \observations(W) \nonumber
+\end{equation}
+This is the register file of the configuration after the first step of execution \ref{eq:ftlr:conf-steps}, so we know:
+\begin{equation}
+  (reg[r_0 \mapsto c, \pcreg \mapsto (\rwx,b',e',a')],h) \step[i-1] (\halted, h') \nonumber
+\end{equation}
+using this and that the register file is in the observation relation, we get a $W' \future W$ and $\hs'$ such that 
+\begin{align*}
+  &h' = \hs' \uplus h_f \\
+  &\heapSat[\hs']{n'-1-(i-1)}{W'}
+\end{align*}
+$W'$ and $\hs'$ satisfy exactly the properties, we need, so we are done.
+
+\item[Case $\reg'(r) = (\exec,b',e',a')$:] This case goes just like the one above with the exception that we need $\readCond{n'-1,b',e',a'}$ which we get in a similar fashion.
+
+\item[Case $\reg'(r) = (\entry,b',e',a')$:] In this case, we do not appeal to the induction hypothesis, but rather to the $\entryCond{}$.
+If $r=\pcreg$, then we have a contradiction with either \ref{eq:ftlr:rwx-assump} or \ref{eq:ftlr:rx-assump}, so it is safe to assume $r \neq \pcreg$.
+If $r=r_0$, then we rely on assumption~\ref{eq:ftlr:cont-in-stdkr} to get $\npair[n']{\entry,b',e',a'} \in \stdvr(W)$.
+In the case where $r \neq r_0$, we rely on assumption~\ref{eq:ftlr:reg-in-stdrr} to get $\npair[n']{\entry,b',e',a'} \in \stdvr(W)$, so we will have $\entryCond{n',W,b',e',a'}$.
+Using the $\entryCond{}$, we get $\npair[n'-1]{(\exec,b',e',a')}\in \stder(W)$.
+From here on, the argument follows like the cases, we have seen.
+One notable thing is that in this case is that $\updatePcPerm{\reg[r_0 \mapsto c, \pcreg \mapsto \pc](r)}$ updates the permission to an execute permission $\exec$ which makes everything align up nicely as in the previous cases.
+
+\item[Otherwise] in this case, we consider the next execution step which goes to $\failed$ as the $\pcreg$ is not a well-formed execution.
+We have, however, assumed that a halting configuration is reached eventually, so a contradiction is reached.
+
+\end{description}
+
+\case{\jnz{}{}}
+\lau{TODO, I have not checked the details as I expect it to be like the $\jmp{}$ case.}
+
+\case{\load{r_1}{r_2}}
+Assume for some $p'$, $b'$, $e'$, $a'$, and $w$ that
+\begin{align}
+  &\reg'(r_1) = (p',b',e',a') \nonumber\\
+  &\readAllowed{p'} \label{eq:ftlr:load-read-allowed} \\
+  &\withinBounds{(p',b',e',a')} \nonumber \\
+  &w = h(a) \nonumber 
+\end{align}
+if the above where not the case, then by the semantics, we would step to a fail configuration which by assumption is not possible. We consider two cases for $r_1$.
+\begin{description}
+\item[$r_1 = \pcreg$] The first step of the execution is
+\[
+  \Phi \step \stdUpdatePc{\updateReg{\pcreg}{w}}
+\]
+If $w$ is not a capability, then
+\[
+\stdUpdatePc{\updateReg{\pcreg}{w}} = \failed
+\]
+and we are done. Assume $w = (p',b',e',a')$, then
+\[
+\stdUpdatePc{\updateReg{\pcreg}{w}} = \updateReg{\pcreg}{(p',b',e',a'+1)}
+\]
+Using execution~\ref{eq:ftlr:conf-steps}, we can conclude that $p' \in \{\rwx,\exec\}$ as the above configuration otherwise would step to $\failed$. From \ref{eq:ftlr:load-read-allowed}
+
+\item[$r_1 \neq \pcreg$]
+\end{description}
+
+\case{\store{}{}}
+
+\case{\move{}{}}
+
+\case{\lea{}{}}
+
+\case{\restrict{}{}{}}
+
+\case{\plus{}{}{}}
+
+\end{proof}
 
@@
   \begin{align*}
-    & \forall W, W', W'' \ldotp W'' \future W' \land W' \future W \Rightarrow W'' \future W
+    & \forall W, W', W'' \ldotp \\
+    & \quad  W'' \future W' \land W' \future W \Rightarrow W'' \future W
   \end{align*}
@@
 \end{align*}
-These are exactly the properties neede to get $W'(r) \nsubeq \iota_{b',e'} (r)$ which is exactly the result we want.
+These are exactly the properties needed to get $W'(r) \nsubeq \iota_{b',e'} (r)$ which is exactly the result we want.
 \end{proof}
@@
 \begin{proof}
-  Follows from uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond$ in the worlds.
+  Follows from uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$ in the worlds.
 That is Lemma~\ref{lem:readCond-mono-world},~\ref{lem:writeCond-mono-world},~\ref{lem:execCond-mono-world}, and~\ref{lem:entryCond-mono-world}.
@@
 Let $\heap_f$, $heap'$, $i \leq n'$ be given and assume $(\reg, \hs \uplus \heap_f) \step[i] (\halted, \heap')$.
-By assumption, we have a $W' \future W$ and $hs'$ such that
+By assumption, we have a $W' \future W$ and $\hs'$ such that
   \begin{align}
-&heap' = hs' \uplus \heap_f \label{lem:eq:stdvr1}\\
-&\heapSat[hs']{n-i}{W'}\label{lem:eq:stdvr2}
+&\heap' = \hs' \uplus \heap_f \label{lem:eq:stdvr1}\\
+&\heapSat[\hs']{n-i}{W'}\label{lem:eq:stdvr2}
   \end{align}. 
-Using $W'$ and $hs'$ as existential witnesses, we already have Equation~\ref{lem:eq:stdvr1} as the first necessary condition and from the above heap satisfaction along with heap satisfaction being uniform in $n$, we get $\heapSat[\hs']{n'-i}{W'}$.
+Using $W'$ and $\hs'$ as existential witnesses, we already have Equation~\ref{lem:eq:stdvr1} as the first necessary condition and from the above heap satisfaction along with heap satisfaction being uniform in $n$, we get $\heapSat[\hs']{n'-i}{W'}$.
 These are the two conditions necessary to get $\npair[n']{(\reg,\hs)} \in \observations(W)$.

### 2016-09-06 capability-machine.tex
Some notes on OS-less stack and return pointer handling.

@@
 
+\section{Other examples and applications}
+\label{sec:other_apps}
+This section contains some ideas about other examples and applications than the
+ticket dispenser example.
+
+\subsection{Stack and return pointer handling without OS involvement using local
+  capabilities}
+The idea of this example would be to work out and prove a calling convention
+that enforces well-bracketed control flow and encapsulation of local variables
+using CHERI's local capabilities.
+
+When one function invokes another function, the essential idea is that:
+\begin{itemize}
+\item Stack pointer is passed as a local and store-local capability.
+\item Return pointer is passed as a local capability.
+\end{itemize}
+
+Since local pointers cannot leave the registers except into regions for which a
+store-local capability is available, this basic idea seems to enforce a number
+of useful properties: well-bracketedness of control flow and encapsulation of
+private state stored on the stack. On the other hand, it also seems to validate
+the standard C treatment of the stack: the stack can be reused after a function
+returns, even between distrusting parties. However, safety/security of this
+design is very non-trivial and seems to rely on some non-trivial reasoning:
+
+\paragraph{Only stack is store-local?}
+A critical assumption is that adversary code has no way to \emph{store} local
+capabilities except on the stack. The reason that it is fine to store local
+capabilities on the stack is that the adversary only has a \emph{local}
+capability to the stack and cannot usefully store that capability anywhere.
+However, this means that we need to rely on the runtime system of our
+programming language to be careful when handing out store-local capabilities:
+only the libc startup code should initialize the stack as store-local and malloc
+should \emph{not} produce them. This basically means that the libc
+initialization code (or whatever component produces the initial stack pointer)
+is part of our TCB.
+
+\paragraph{Requirement for clearing the stack}
+Imagine the following trusted C function:
+
+\begin{verbatim}
+void myfunction(){
+  advfunction1();
+  advfunction2();
+}
+\end{verbatim}
+
+where advfunction1() and advfunction2() are adversary functions. In the standard
+C treatment of the stack, advfunction2() would get the same stack pointer as
+advfunction1(). This is supposed to be safe since advfunction1() must have
+released all capabilities to the stack after its execution. But what if we
+require that the two functions have no way of communicating with each other?
+Concretely, advfunction1() has access to some secrets that must not be leaked to
+advfunction2(). How can we prevent advfunction1() from storing the secret
+somewhere on the stack and relying on advfunction2() from receiving the same
+stack pointer where it can read the secret? The most obvious solution seems to
+be that we should fully clear the stack (overwrite it with zeros) after the
+return of any adversary function, but this could cause an important overhead.
+Perhaps the processor should accomodate this with a special instruction that can
+zero the entire array that a capability points to?
+
+\paragraph{What do return pointers look like?}
+An important question is what return pointers look like? Since we want to
+protect the caller from the callee, it's important that the return pointer is
+opaque, i.e. an entry pointer. The entry pointer will point to a closure that
+contains the next instruction to execute, as well as the previous stack pointer.
+But since stack pointers are local, this means that the return pointer closure
+should be stored in a region of memory for which we have store-local permission,
+i.e. on the stack. This means we need the following in our calling convention:
+before invoking a function, we push the stack pointer and the instruction
+pointer after invocation on the stack, we construct a return pointer by
+copying the stack pointer, limiting it to these two entries and making it an
+entry pointer.  Then we shrink the stack pointer to the unused part of the stack
+and jump.
+
+\paragraph{Only one-way protection in higher-order settings?}
+Another important point is that, in a sense, local capabilities provide only
+one-way protection: the caller is protected from the callee but not vice-versa.
+Concretely: when invoking a function with some arguments marked as local, the
+caller is guaranteed that the callee will not have been able to store the
+capabilities anywhere (except perhaps on the stack, see above). However, the
+callee seems to have more limited guarantees: Particularly, the caller may have
+kept its own stack capability and this stack capability may (and typically will)
+also cover the part of the stack that is ``owned'' by the callee.  In this
+sense, the guarantees are more limited than in a linear language.
+
+So what does this mean? In a first-order language, this is all fine, but what if
+we are in a higher-order language. Imagine the following (in some ML-like language):
+
+\begin{verbatim}
+let f = fun callback =>
+          let ... in 
+          let ret = callback() in
+          ...
+//adversary top function
+let advtop = f( (fun y => ...) )
+\end{verbatim}
+
+Our trusted function f is invoked by the adversary (from function advtop()) and
+wants to be able to invoke an untrusted callback received from the adversary.
+When invoking the closure, we don't want it to be able to access f's local
+variables which it has stored on the stack. To achieve this, we only give it a
+stack pointer that covers the part of the stack that is unused by f. However,
+the callback may be implemented as an entry pointer that carries capabilities,
+particularly the capability to advtop's stack pointer, which includes the part
+of the stack that is now used by f and contains f's local variables.
+
+So how do we deal with this? Perhaps we should use the fact that this is only
+possible when f's callback argument is allocated to some part of the memory to
+which advtop has store-local permissions (since the callback contains a
+reference to the stack to which advtop only has a local capability). One way to
+exclude the scenario is to require that callbacks are provided as non-local
+capabilities. The downside of this is that local callbacks can be useful for the
+caller to prevent the callee from storing them. Another way to exclude the
+scenario is to require that the stack is allocated in a fixed part of the
+address space and to check that callbacks point outside of this region before
+invoking them.
+
 \section{Related reading}

### 2016-09-09 capability-machine.tex
Fix typo.

@@
 \begin{align*}
-  &W'(r).\phi = W(r).phi \\
+  &W'(r).\phi = W(r).\phi \\
   &(W'(r).s,W(r).s) \in W(r).\phi \\
@@
 Our trusted function f is invoked by the adversary (from function advtop()) and
-wants to be able to invoke an untrusted callback received from the adversary.
-When invoking the closure, we don't want it to be able to access f's local
-variables which it has stored on the stack. To achieve this, we only give it a
-stack pointer that covers the part of the stack that is unused by f. However,
-the callback may be implemented as an entry pointer that carries capabilities,
-particularly the capability to advtop's stack pointer, which includes the part
-of the stack that is now used by f and contains f's local variables.
+wants to invoke an untrusted callback received from the adversary. When invoking
+the closure, we don't want it to be able to access f's local variables which it
+has stored on the stack. To achieve this, we only give it a stack pointer that
+covers the part of the stack that is unused by f. However, the callback may be
+implemented as an entry pointer that carries capabilities, particularly the
+capability to advtop's stack pointer, which includes the part of the stack that
+is now used by f and contains f's local variables.
 
@@
 which advtop has store-local permissions (since the callback contains a
-reference to the stack to which advtop only has a local capability). One way to
-exclude the scenario is to require that callbacks are provided as non-local
-capabilities. The downside of this is that local callbacks can be useful for the
-caller to prevent the callee from storing them. Another way to exclude the
-scenario is to require that the stack is allocated in a fixed part of the
-address space and to check that callbacks point outside of this region before
-invoking them.
+reference to the stack to which advtop only has a local capability). I see
+basically three ways to do this, all based on the idea of enforcing that the
+callback should be constructed in a part of memory for which no store-local
+permissions are available:
+\begin{itemize}
+\item One way to exclude the scenario is to require that callbacks are provided
+  as non-local capabilities. The downside of this is that local callbacks can be
+  useful for the caller to prevent the callee from storing them.
+\item Another way to exclude the scenario is to require that the stack is
+  allocated in a fixed part of the address space and to check that callbacks
+  point outside of this region before invoking them.
+\item Perhaps we should require that store-local permissions cannot be removed
+  from a capability and simply require that callback pointers do not have
+  store-local set. Perhaps we can allow store-local permissions to be given up,
+  but only if the corresponding part of memory is fully zeroed in the process
+  (or at least all local capabilities stored in the region).
+\end{itemize}
 

### 2016-09-13 capability-machine.tex
Reformulation to prevent confusion.
A simple result to prove about the OS-less stack and return pointer passing example.

@@
 C treatment of the stack, advfunction2() would get the same stack pointer as
-advfunction1(). This is supposed to be safe since advfunction1() must have
-released all capabilities to the stack after its execution. But what if we
-require that the two functions have no way of communicating with each other?
-Concretely, advfunction1() has access to some secrets that must not be leaked to
+advfunction1(). This is supposed to be safe since advfunction1() cannot have
+kept capabilities for the stack after its execution. But what if we require that
+the two functions have no way of communicating with each other? Concretely,
+advfunction1() has access to some secrets that must not be leaked to
 advfunction2(). How can we prevent advfunction1() from storing the secret
@@
 
+\subsection{A result to prove...}
+\label{sec:os-less-stack-property}
+
+The simplest thing that comes to mind as a formal result for all of the above is
+to look at a concrete program that clearly relies on properties like
+well-bracketed control flow and encapsulation of local variables and prove it
+correct. As a concrete example: we might show an assembly program that
+corresponds to the following (a higher-order program that crosses trust
+boundaries and relies on local variable encapsulation and well-bracketed control
+flow):
+\begin{verbatim}
+let trustedCode = fun adversary =>
+                    let x = ref 0 in
+                    let callback = fun adv2 => let y = ref (!x) in assert (!x >= 0); x := -1; c2 unit; x := !y)
+                    let _ = adversary callback
+                    assert (!x >= 0)
+\end{verbatim}
+
 \section{Related reading}

### 2016-09-19 capability-machine.tex
Added subsection on worlds and appendix with some of the proofs

--- 2016-09-06 capability-machine.tex
@@
 
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
+\usepackage{amsmath,amsfonts,amssymb,amsthm}
 \usepackage{thmtools}
@@
 \usepackage{ifpdf}
+\usepackage{verbatim}
+
+\usepackage{mathtools}
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
+
+\usepackage{appendix}
 
@@
 \newcommand{\parfun}{\rightharpoonup}
+\newcommand{\finparfun}{\stackrel{\textit{\tiny{fin}}}{\rightharpoonup}}
 \newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\longrightarrow}}
 \newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
+\newcommand{\nefun}{\stackrel{\textit{\tiny{ne}}}{\longrightarrow}}
+\newcommand{\fun}{\longrightarrow}
 \newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
@@
 \DeclareMathOperator{\dom}{dom}
-
+\newcommand{\blater}{\mathop{\blacktriangleright}}
+\newcommand{\id}{\var{id}}
 \newcommand{\undefined}{\mathit{undefined}}
+
+\newcommand{\powerset}[1]{\mathcal{P}(#1)}
 
@@
 \newcommand{\true}{\mathit{true}}
+
+%cofes
+\newcommand{\cofe}{c.o.f.e.}
+\newcommand{\cofes}{\cofe{}'s}
+\newcommand{\CatC}{\mathbb{C}}
+\newcommand{\CatP}{\mathbb{P}}
 
@@
 
+\newcommand{\Rel}{\plaindom{Rel}}
 \newcommand{\States}{\plaindom{State}}
@@
 \label{sec:logical-relation}
+\subsection{Recursive Domain Equation}
+\label{subsec:recursive-dom-eq}
+The goal is to solve the following domain equation:
+\[
+W = \nats \finparfun (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments})))
+\]
+Where $\States$ is a set of states with all the ones we use in this paper.
+\[
+\Rel= \{R \in \powerset{\States^2} \mid R \text{ is reflexive and transitive} \}
+\]
+
+This cannot be solved with sets, so we use preordered complete ordered families of equivalences where it is possible to solve such an equation that ressembles the above one, namely it is possible to find an isomorphism $\xi$ and preordered \cofe{} $W$ such that
+\[
+  \xi : W \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments}))))
+\]
+
+\begin{definition}[o.f.e's]
+  An ordered family of equivalences (o.f.e.) is a set and a family of equivalences, $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$.
+The familly of equivalences have to satisfy the following properties
+  \begin{itemize}
+  \item $\nequal[0]$ is a total relation on $X$
+  \item $\forall n \ldotp \forall x,y \in S \ldotp x \nequal[n+1] y \Rightarrow x \nequal y$
+  \item $\forall x,y \ldotp(\forall n \ldotp x \nequal y) \Rightarrow x=y$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}[\cofes]
+  A complete orderede family of equivalences is an o.f.e.\ $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ where all Cauchy sequences in $X$ have a limit in $X$.
+\end{definition}
+
+\begin{definition}[Preordered \cofes]
+  A preordered \cofe{}\ is a \cofe{}\ equiped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. That is the ordering must satisfy the following
+  \begin{itemize}
+  \item $\forall x \in X \ldotp x \future x$
+  \item $\forall x,y,z \in X \ldotp x \future y \land y \future z \Rightarrow x \future z$
+  \end{itemize}
+\end{definition}
+
+The category of \cofes{} is the category whith \cofes{} as objects and non-expansive functions as morphisms.
+We denote this category $\CatC$.
+The category of preordered \cofes{} has preordered \cofes as objects and monotone and non-expansive functions as morphisms.
+We denote this category $\CatP$.
+
+\begin{definition}[Preordered \cofe{}\ construction: Finite-partial function]
+  Given a set $S$ and preordered \cofe{}\ $X$, $S \finparfun X$ is a preordered \cofe{}\ with the ordering
+  \[
+    \begin{gathered}
+      f \future g  \\
+      \text{ iff } \\
+      \dom(f) \supseteq \dom(g) \; \text{ and } \; \forall n \in S \ldotp f(n) \future g(n)
+    \end{gathered}
+  \]
+\end{definition}
+
+We need the following constructions to create the preordered \cofe{}, we want to work use
+\begin{definition}[Preordered \cofe{}\ construction: Function]
+  Given a set $S$ and \cofe{}\ $\var{HP}$, $S \fun \var{HP}$ is a preordered \cofe{}\ with the ordering
+\[
+  \begin{gathered}
+    f \future g \\
+    \text{ iff } \\
+    \forall s \in \dom(f) \ldotp f(s) \future g(s)
+  \end{gathered}
+\]
+\end{definition}
+
+\begin{definition}[Preordered \cofe{}\ construction: Monotone, non-expansive function]
+  Given a preordered \cofe{}\ $W$ and preordered \cofe{}\ $U$, $W \monnefun U$ is a preordered \cofe{} with the ordering
+\[
+  \begin{gathered}
+    f \future g \\
+    \text{ iff } \\
+    \forall s \in \dom(f) \ldotp f(s) \future g(s)
+  \end{gathered}
+\]
+\end{definition}
+
+The above are standard constructions, so they are used here without showing they infact are well-defined.
+
+\begin{definition}[Preordered \cofe{}\ construction: Region]
+  \label{def:pcofe-region}
+  Given a \cofe{}\ $H$, the tuple 
+\[
+  (\States \times \Rel \times H)
+\]
+ is a preordered \cofe{}\ with the ordering
+\[
+  \begin{gathered}
+    (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_2)\\
+    \text{ iff }\\
+    H_2 = H_1 \text{ and } \phi_2 = \phi_1 \text{ and } (s_1,s_2) \in \phi_2\\
+  \end{gathered}
+\]
+\end{definition}
+\begin{lemma}[Region definition well-defined]
+\label{lem:reg-def-wd}
+  The construction in Definition~\ref{def:pcofe-region} is a preordered \cofe{}. That is
+  \begin{itemize}
+  \item It is a \cofe{} (this is a standard construction)
+  \item $\future$ is a transitive and reflexive relation.
+  \item $\future$ preserves limits. \\That is for Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$ if 
+    \[
+      \{a_n\}_n \geq \{b_n\}_n,
+    \]
+    then 
+    \[
+      \lim \{a_n\}_n \future \lim \{b_n\}_n
+    \]
+  \end{itemize}
+\end{lemma}
+
+Define functors $K$, $R$, and $G$ as follows:
+\begin{align*}
+  & K : \CatP \rightarrow \CatP \\
+  & K(R) = \nats \finparfun R \\
+  & K(f) = \lambda \phi \ldotp \lambda n \ldotp f(\phi(n)) \\
+  & \\
+  & R : \CatC \rightarrow \CatP \\
+  & R(H) = \States \times \Rel \times H\\
+  & R(h) = \lambda (s,\Phi,H) \ldotp (s,\Phi,h(H))\\
+  & \\
+  & G : \CatP^{\var{op}} \rightarrow \CatC \\
+  & G(W) = \UPred{HS}^{W^{\States}} \\
+  & G(g) = \lambda H \ldotp \lambda \var{st} \ldotp \lambda x \ldotp H (\var{st})(g(x))
+\end{align*}
+We first show that $K$, $R$, and $G$ are well-defined mappings.
+\begin{lemma}[World finite partial mapping]
+  \label{lem:world-finpar}
+  For all $f$ and $\phi$, $K(f)(\phi)$ is a finite partial mapping.
+\end{lemma}
+\begin{lemma}[Heap segment predicate monotone]
+\label{lem:HSP-mon}
+For all $g$, $H$, and $\var{st}$
+\[
+G(g) (H) (\var{st})
+\]
+is non-expansive.
+\end{lemma}
+\begin{lemma}[Heap segment predicate non-expansive]
+\label{lem:HSP-ne}
+For all $g$, $H$, and $\var{st}$
+\[
+G(g) (H) (\var{st})
+\]
+is monotone.
+\end{lemma}
+Next we show that $K$, $R$, and $G$ are in fact functors:
+\begin{lemma}[$K$ functorial]\forcenewline
+\label{lem:K-func}
+  \begin{enumerate}
+  \item $K(f) : K(X) \rightarrow K(Y)$ is monotone and non-expansive for $f : X \monnefun Y$
+  \item $K(f \circ g) = K(f) \circ K(g)$ for $f : Z \nefun Y$ and $g : X \nefun Z$
+  \item $K(\id) = \id$
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}[$R$ functorial]\forcenewline
+\label{lem:R-func}
+  \begin{enumerate}
+  \item $R(f) : R(X) \rightarrow R(Y)$ is non-expansive and monotone for $f : X \nefun Y$
+  \item $R(f \circ g) = R(f) \circ R(g)$ for $f : Z \nefun Y$ and $g : X \nefun Z$
+  \item $R(\id) = \id$
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}[$G$ functorial]\forcenewline
+\label{lem:G-func}
+  \begin{enumerate}
+  \item $G(f) : G(Y) \rightarrow G(X)$ is non-expansive for $f : X \monnefun Y$
+  \item $G(f \circ g) = G(g) \circ G(f)$ for $f : Z \nefun Y$ and $g : Y \nefun Z$
+  \item $G(\id) = \id$
+  \end{enumerate}
+\end{lemma}
+We now compose the above functors into the functor we actually want to use: $F = K \circ R \circ G$, $F : \CatP^{\var{op}} \rightarrow \CatP$.
+\begin{lemma}[$F$ functorial]\forcenewline
+\label{lem:F-func}
+  \begin{enumerate}
+  \item $F(f) : F(Y) \rightarrow F(X)$ is monotone and non-expansive for $f : X \monnefun Y$
+  \item $F(f \circ g) = F(g) \circ F(f)$ for $f : Z \nefun Y$ and $g : Y \nefun Z$\
+  \item $F(\id) = \id$
+  \end{enumerate}
+\end{lemma}
+
+\begin{lemma}[$F$ locally non-expansive]
+\label{lem:F-loc-ne}
+For all $f, g : X \fun Y$, if $f \nequal g$, then $F(f) \nequal F(g)$.
+\end{lemma}
+With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-copmosition of $blater$:
+\[
+  F'(W) = \blater (F(W))
+\]
+We have a theorem that gives us a solution to the recurisve domain equation
+\[
+  W \cong F'(W) = \blater (\nats \finparfun (\States \times \Rel \times (\States \fun W \monnefun \UPred{\HeapSegments} )))
+\]
+The solution to the recursice domain equations is presented \cite{Birkedal:2010:TCS:411:4102-4122}.
+They solve it in pre-ordered, non-empty, complete, 1-bounded ultrametric spaces, but they have a simple correspondence to pre-ordered \cofes{}.
+
 \subsection{Worlds}
-\[
-  \Worlds = \{ W \in RegionNames \parfun \Regions \mid \dom(W) \text{ finite } \}
-\]
-$\States$ is a fixed set with the necessary states.
-
-\begin{align*}
-  \Regions = \{ r = (s, \phi, H) \mid &  s \in \States \land \phi \subseteq \States^2 \land \\ 
-                                      &  H : \States \rightarrow \StorePred \land \\ 
-                                      &  \phi \text{ reflexive and transitive} \}
-\end{align*}
-For convenience, we write $r.s$, $s.\phi$, and $s.H$ for a the projection of a regions state, transition system, and store predicate, respectively.
-\[
-  \StorePred = \widehat{\Worlds} \monnefun \UPred{\HeapSegments}
-\]
-
-\[
-\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
-\]
-Future Worlds:
-\[
- W' \future W \text{ iff }  
+Assume preordered \cofe{}\ $W$ and isomorphism $\xi$ such that:
+\[
+  \xi : W \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments}))))
+\]
+We now define regions as
+\[
+\Regions \defeq (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments})))
+\]
+and we define worlds as
+\[
+\Worlds \defeq \RegionNames \finparfun \Regions
+\]
+To define future worlds and regions, We use the ordering inherited from the preordered \cofes{}.
+\begin{definition}[Future worlds]
+  For $W, W' \in \Worlds$
+ \begin{align*}
+ W' \future W & & \text{ iff } & &
    \begin{gathered}
@@
    \end{gathered}
-\]
-Future regions:
-\[
-  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
-    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2) \in \phi_1
-\]
-n-subset for regions:
-\[
-(s_1,\phi_1,H_1) \nsubeq (s_2,\phi_2,H_2) \text{ iff } 
+ \end{align*}
+\end{definition}
+
+\begin{definition}[Future regions]
+For regions $(s_2,\phi_2,H_2), (s_1,\phi_1,H_1) \in \Regions$
+  \begin{align*}
+ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) &&\text{ iff } & &
+(s_1,\phi_1) = (s_2,\phi_2) \text{ and } (s_1,s_2) \in \phi_2
+  \end{align*}
+\end{definition}
+
+\begin{definition}[$n$-subset for regions]
+For regions $(s_1,\phi_1,H_1), (s_2,\phi_2,H_2) \in \Regions$
+\begin{align*}
+  (s_1,\phi_1,H_1) \nsubeq (s_2,\phi_2,H_2) & & \text{ iff } & &
   \begin{gathered}
@@
     \text{ and }\\
- \forall \hat{W} \ldotp H_1 \; s_1 \; \hat{W} \nsubeq H_2 \; s_2 \; \hat{W}
+    \forall \hat{W} \in \blater \Worlds \ldotp H_1 \; s_1 \; \hat{W} \nsubeq H_2 \; s_2 \; \hat{W}
   \end{gathered}
-\]
+\end{align*}
+\end{definition}
+
 \begin{definition}[Heap satisfaction/erasure]
-\[
-\heapSat[\hs]{n}{W} \text{ iff }
+\begin{align*}
+&\heapSat[\hs]{n}{W} \\
+&\qquad\text{ iff }\\
+&\exists R : \dom(W) \rightarrow \HeapSegments \ldotp & &\\
+&\quad\begin{aligned}
+  &\hs = \biguplus_{r \in \dom(W)} R(r) \\
+  &\qquad\text{and}\\
+  &\forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\xi^{-1}(W))
+\end{aligned} 
+\end{align*}
+
+\end{definition}
+
+\subsection{Standard Regions}
+\label{subsec:standard-regions}
+A standard heap invariant that insures all values in the region are in the value  relation:
+\[
+  \iota_{\var{start},\var{end}} \; : \; \Regions
+\]
+\begin{align*}
+  \iota_{\var{base},\var{end}} & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
+  H_{\var{std}}\; (\var{base},\var{end})\; W &  \defeq \left\{ \npair{\hs} \middle|
                                                \begin{aligned}
-&\exists R : \dom(W) \rightarrow \HeapSegments \ldotp \\
-&\quad\begin{gathered}
-  \hs = \biguplus_{r \in \dom(W)} R(r) \\
-  \text{and}\\
-  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
-\end{gathered}
-\end{aligned}
-\]
-\end{definition}
-
-A standard heap invariant that insures all values in the region are in the value  relation:
-\[
-  \iota_{\var{start},\var{end}} \; : \; \Regions
-\]
-\begin{align*}
-  \iota_{\var{base},\var{end}} & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
-  H_{\var{std}}\; (\var{base},\var{end})\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\roll \; W) \\
-                             & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
-\end{align*}
+                                                 & \dom(\hs) = [\var{base},\var{end}] \land \\
+                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; W)
+                                               \end{aligned} \right\}
+\end{align*}
+\begin{lemma}[$\iota_{\var{start},\var{end}}$ is well-defined]
+\label{lem:std-region-wd}
+  For all $\var{base}$ and $\var{end}$, $H_{\var{std}}\; (\var{base}, \var{end})$ is monotone and non-expansive. $=$ is a reflexive and transitive relation.
+\end{lemma}
 
@@
   \iota_{\var{base},\var{end}}^0 & \defeq ((\var{base},\var{end}),=,H_{\var{std}}^0) \\
-  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{base},\var{end}] \land \\
-                                                                & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
-\end{align*}
-
+  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\left\{ \npair{\heapseg} \middle| 
+                                                     \begin{aligned}
+                                                       & \dom(\heapseg) = [\var{base},\var{end}] \land \\
+                                                       & \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  
+                                                     \end{aligned} \right\}
+\end{align*}
+
+\begin{lemma}[$\iota_{\var{base},\var{end}}^0$ is well-defined]
+\label{lem:std-z-region-wd}
+    For all $\var{base}$ and $\var{end}$, $H_{\var{std}}^0 \; (\var{base},\var{end})$ is monotone and non-expansive. $=$ is a reflexive and transitive relation.
+\end{lemma}
+
+\subsection{Value Relation}
+In order to define the value relation consiely using the standard regions from the Subsection~\ref{subsec:standard-regions}, we define the following conditions:
 \begin{align*}
@@
                                           & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp\\
-                                          & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'}(r) \\ & \\
+                                          & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
   \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionNames \ldotp \\
@@
 
+\begin{lemma}[Heap satisfaction non-expansive]
+  \begin{align*}
+    & \forall n,\heap, W, W' \ldotp\\
+    & \quad \heapSat{n}{W} \land W \nequal W' \Rightarrow \heapSat{n}{W'}
+  \end{align*}
+\end{lemma}
+
 \begin{lemma}[Read condition uniformity]
@@
 
+\begin{appendices}
+\section{Definitoins and Proofs related to Section~\ref{subsec:recursive-dom-eq}}
+\end{appendices}
+
+\begin{definition}[Ordering of Cauchy chains]
+  For Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$,
+  \begin{align*}
+    \{a_n\}_n \geq \{b_n\}_n & & \text{iff} & & \forall n \ldotp a_n \geq b_n
+  \end{align*}
+\end{definition}
+
+\begin{definition}[Limit of Cauchy chains]
+  For a Cauchy chain $\{a_n\}_n$ in the \cofe{}\ $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$, we say that $a$ is a limit  (denoted $\lim\{a_n\}_n$) of this chain iff it satisfies
+  \begin{align*}
+    & \forall n \ldotp \exists N \ldotp \forall m \geq N \ldotp a \nequal a_m
+  \end{align*}
+\end{definition}
+
+\begin{definition}[Set \cofe{}\ Construction]
+  Given a set $X$, then $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ is a \cofe{}\ where all the equivalences are equiality.
+\end{definition}
+
+\begin{definition}[Function \cofe{}\ Construction]
+  Given \cofes{} $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the \cofe{}\ $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) \fun \left(X, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ as
+\[
+\left(\{f : X \fun Y \mid f \text{ non-expansive}\}, \left(\nequal\right)_{n=0}^{\infty} \right)
+\]
+where the equivalence is defined as
+\begin{align*}
+  f \nequal g & & \text{ iff } & & \forall x \in X \ldotp f(x) \nequal_Y g(x)
+\end{align*}
+
+\end{definition}
+
+\begin{proof}[Proof of Lemma~\ref{lem:reg-def-wd}]
+  The \cofe{}\ is a standard construction for tuples, so we won't show that it in fact is a \cofe{}\ here.
+That $\future$ is reflexive and transitive follows trivially from $\Phi$ being reflexive and transitive.
+It remains to show that $\future$ preserves limits.
+
+  Assume $\{a_n\}_n \geq \{b_n\}_n$.From this it follows that for all $m \in \nats$:
+  \begin{align*}
+    \Phi_{b_m} & = \Phi_{a_m} \\
+    H_{b_m} & = H_{a_m} \\
+    (s_{a_m},s_{b_m}) &\in \Phi_{b_m}
+  \end{align*}
+  Say $\lim{a_n}_n = (s_a,\Phi_a,H_a)$ and $\lim{b_n}_n = (s_b,\Phi_b, H_b)$. We need to show
+  \begin{align*}
+    \Phi_{b} & = \Phi_{a} \\
+    H_{b} & = H_{a} \\
+    (s_{a},s_{b}) &\in \Phi_{b}
+  \end{align*}
+  To show $H_{b} = H_{a}$ by one of the \cofe{}\ properties SFTS $\forall n \ldotp H_{b} \nequal H_{a}$. Given $n$ use that $a$ is a limit to get $N$ s.t.
+\[
+  \forall m \geq N \ldotp H_{a_m} \nequal H_a
+\]
+Likewise for $b$ we get a similar $M$. Take some $m \geq \max\{M,N\}$, using $H_{b_m} = H_{a_m}$, we get:
+\[
+  H_a = H_{a_m} = H_{b_m} = H_b
+\]
+
+For $\Phi_{b} = \Phi_{a}$ we do something similar except, we just pick $n=1$ as the $n$-equality in this \cofe{} is equality. 
+Same approach is used for $(s_{a},s_{b}) \in \Phi_{b}$
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:world-finpar}]
+\[
+  K(f)(\phi) = \lambda n \ldotp f(\phi(n))
+\]
+The above is defined exactly on the $n\in \nats$ where $\phi$ is defined. $\phi$ is a finite partial function, so the above is as well.
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:HSP-mon}]
+  Given $g$, $H$, $\var{st}$, $w$ and $w'$. Assume $w' \future w$. Show
+  \begin{gather}
+    (G(g))(\phi)(\var{st})(w') \future (G(g))(\phi)(\var{st})(w) \nonumber\\ 
+    \Uparrow \label{lem:HSP-mon:b1}\\
+    H(\var{st})(g(w')) \future H(\var{st})(g(w)) \nonumber \\
+    \Uparrow \label{lem:HSP-mon:b2}\\
+    g(w') \future g(w) \nonumber\\
+    \Uparrow \label{lem:HSP-mon:b3}\\
+    w' \future w \nonumber
+  \end{gather}
+(\ref{lem:HSP-mon:b1}) expands the definition of $G$.
+(\ref{lem:HSP-mon:b2}) is due to $H(\var{st})$ being monotone.
+(\ref{lem:HSP-mon:b3}) is due to $g$ being a morphism in $\CatC^{\var{op}}$ and they are monotone function.
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:HSP-ne}]
+  Given $n$, $g$, $H$, $\var{st}$, $w$ and $w'$. Assume $w' \nequal w$. Show
+  \begin{gather}
+    (G(g))(\phi)(\var{st})(w') \nequal (G(g))(\phi)(\var{st})(w) \nonumber\\ 
+    \Uparrow \nonumber \\
+    H(\var{st})(g(w')) \nequal H(\var{st})(g(w)) \nonumber \\
+  \end{gather}
+  We use that $H=H$ implies $H \nequal H$ which means
+  \[
+    \forall \var{st}, \var{st'} \ldotp \var{st} \nequal \var{st'} \Rightarrow H(\var{st}) \nequal H(\var{st'})
+  \]
+  This in turn implies
+  \[
+    \forall y,y' \ldotp \ldotp y \nequal y' \Rightarrow H(\var{st})(y) \nequal H(\var{st})(y')
+  \]
+  Using our assumption, about the worlds, we are done.
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:K-func}]
+  $K(f) : K(X) \rightarrow K(Y)$ is \emph{monotone} for $f : X \monnefun Y$:
+  Assume $\phi \future \phi'$. Show
+  \[
+    (K(f))(\phi) \future (K(f))(\phi') 
+  \]
+  By def on ordering for functions SFTS
+  \[
+    \forall n \ldotp (K(f))(\phi)(n) \future (K(f))(\phi')(n)
+  \]
+  let $n$ be givenm show
+  \[
+    f(\phi'(n)) \future f(\phi(n))
+  \]
+  $f$ mon, so SFTS
+  \[
+    \phi'(n) \future \phi(n)
+  \]
+  which follows from the assumption $\phi' \future \phi$.\\\\
+  $K(f) : K(X) \rightarrow K(Y)$ is \emph{non-expansive} for $f : X \monnefun Y$:
+  Assume $\phi \nequal \phi'$. Show
+  \[
+    (K(f))(\phi) \nequal (K(f))(\phi') 
+  \]
+  Let $k$ be given, SFTS
+  \[
+    (K(f))(\phi)(k) \nequal (K(f))(\phi')(k)
+  \]
+  which unfolds to
+  \[
+    f(\phi(k)) \nequal f(\phi'(k))
+  \]
+  as $f$ is non-expansive SFTS
+  \[
+    \phi(k) \nequal \phi'(k)
+  \]
+  which follows from assumption.\\\\
+  $K$ preserves composition, show
+  \[
+    K(f \circ g) = K(f) \circ K(g)
+  \]
+  Given $\phi$ and $n$ show
+  \begin{align*}
+    (K(f) \circ K(g))(\phi)(n) & = K(f)(K(g)(\phi))(n) \\
+                               & = f(K(g)(\phi)(n)) \\
+                               & = f(g(\phi(n))) \\
+                               & = (f \circ g)(\phi(n)) \\
+                               & = K(f \circ g)(\phi)(n)
+  \end{align*}\\\\
+  $K$ preserves identity, show
+  \[
+    K(\id) = \id
+  \]
+  Let $\phi$ and $n$ be given.
+  \begin{align*}
+    K(\id) & = \lambda \phi \ldotp \lambda n \ldotp \id(\phi(n)) \\
+           & = \lambda \phi \ldotp \lambda n \ldotp \phi(n) \\
+           & = \lambda \phi \ldotp \phi \\
+           & = \id
+  \end{align*}
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:R-func}]
+   $R(f) : R(X) \rightarrow R(Y)$ is \emph{monotone} for $f : X \nefun Y$. Given aformentioned $f$. Further let $(s_2,\Phi_2,H_2) \future (s_1,\Phi_1,H_1)$ be given.
+
+    \begin{gather*}
+      R(f)(s_2,\Phi_2,H_2) = (s_2,\Phi_2,H_2(f)) \future (s_1,\Phi_1,H_1(f)) = R(f)(s_1,\Phi_1,H_1) \\
+      \Uparrow\\
+      \Phi_2 = \Phi_1 \text{ and } H_1(f) = H_2(f) \text{ and } (s_1,s_2) \in \Phi_2
+    \end{gather*}
+    From our assumption, we have $H_2 = H_1$ from which it follows $H_1(f) = H_2(f)$. The remaining conditions also follow directly from the assumption.
+
+    $R(f) : R(X) \rightarrow R(Y)$ is \emph{non-expansive} for $f : X \nefun Y$. Let $(s_2,\Phi_2,H_2) \nequal (s_1,\Phi_1,H_1)$ be given
+\[
+  R(f)(s_2,\Phi_2,H_2) = (s_2,\Phi_2,H_2(f)) \nequal (s_1,\Phi_1,H_1(f)) = R(f)(s_1,\Phi_1,H_1)
+\]
+$H_1(f) \nequal H_2(f)$ follows from the definition of $n$-equality on functions as well as the fact that $f=f$ implies $f \nequal f$ for any $n$.
+
+  Composition distributes over $R$, $R(f \circ g) = R(f) \circ R(g)$.
+  \begin{align*}
+    R(f) \circ R(g)(s,\Phi,H) &= R(f)(s,\Phi,g(H)) \\
+                              &= (s,\Phi,f(g(H)))\\
+                              &= R(f \circ g)
+  \end{align*}
+
+  Idenntity distributes over $R$, $R(\id) = \id$.
+  \[
+    R(id) = \lambda (s,\Phi,H) \ldotp  (s,\Phi,\id(H)) = \lambda (s,\Phi,H) \ldotp  (s,\Phi,H) = \id
+  \]
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:G-func}]
+  $G(f) : G(Y) \rightarrow G(X)$ is \emph{non-expansive} for $f : X \monnefun Y$. For $H \nequal H'$ show
+  \[
+    G(f)(H) \nequal G(f)(H')
+  \]
+  which expands to
+  \[
+    \lambda \var{st} \ldotp \lambda w \ldotp H(\var{st})(f(w)) \nequal \lambda \var{st} \ldotp \lambda w \ldotp  H'(\var{st})(f(w))
+  \]
+  Given $\var{st} \nequal \var{st}'$ and $w \nequal n'$ SFTS
+  \[
+  \ldotp H(\var{st})(f(w)) \nequal H'(\var{st}')(f(w'))
+  \]
+  As $f$ is non-expansive, we get $f(w) \nequal f(w')$. As $H \nequal H'$ we get $H(\var{st}) \nequal H(\var{st'})$. Combining the two, we get the desired result.  
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:F-func}]
+Follows from Lemmas~\ref{lem:K-func},~\ref{lem:R-func}, and~\ref{lem:G-func}.
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:F-loc-ne}]
+Let $k$, $f$ and $g$ be given and assume $f \nequal[k] g$. Show
+\[
+F(f) \nequal[k] F(g).
+\]
+SFTS
+\[
+\forall \phi \ldotp (F(f))(\phi) \nequal[k] (F(g))(\phi)
+\]
+Let $\phi$ be given. We need to show $\dom((F(f))(\phi)) = \dom((F(g))(\phi))$. Both sides are defined exactly when $\phi$, so they have the same domain. We also need to show
+\[
+\forall n \in \dom((F(f))(\phi))\ldotp (F(f))(\phi)(n) \nequal[k] (F(g))(\phi)(n)
+\]
+Let $n \in  \dom((F(f))(\phi))$ be given. Define $H = \pi_3 (\phi(n))$. The above unfolds to two triples, so SFTS each component is $k$-equal. That is
+\begin{align*}
+  \pi_1(\phi(n)) \nequal[k] \pi_1(\phi(n))\\
+  \pi_2(\phi(n)) \nequal[k] \pi_2(\phi(n))\\
+  \lambda \var{st}'\ldotp (H(\var{st}') \circ f) \nequal[k]   \lambda \var{st}'\ldotp (H(\var{st}') \circ g)
+\end{align*}
+The two first are trivially satisfied. For the third, we need to show that for all $\var{st}$ if we apply $\var{st}$ to each side, then they are still $k$-equal. That is given $\var{st}$ show
+\[
+  (H(\var{st}) \circ f) \nequal[k]   \ldotp (H(\var{st}) \circ g)
+\]
+SFTS
+\[
+\forall w \ldotp (H(\var{st}) \circ f)(w) \nequal[k]   \ldotp (H(\var{st}) \circ g)(w)
+\]
+Let $w$ be given. From $f \nequal[k] g$, we get $f(w) \nequal[k] g(w)$. From this and $H(st)$ being non-expansive, we get
+\[
+  H(\var{st})(f(w)) \nequal[k]   H(\var{st})(g(w))
+\]
+which is what we neede to show.
+
+\end{proof}
+
+\section{Proof of Lemmas in Section~\ref{subsec:standard-regions}}
+\begin{proof}[Proof of Lemma~\ref{lem:std-region-wd}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:std-z-region-wd}]
+  %TODO
+\end{proof}
+
 \end{document}
+\grid

### 2016-09-19 capability-machine.tex

@@
 which advtop has store-local permissions (since the callback contains a
-reference to the stack to which advtop only has a local capability). One way to
-exclude the scenario is to require that callbacks are provided as non-local
-capabilities. The downside of this is that local callbacks can be useful for the
-caller to prevent the callee from storing them. Another way to exclude the
-scenario is to require that the stack is allocated in a fixed part of the
-address space and to check that callbacks point outside of this region before
-invoking them.
+reference to the stack to which advtop only has a local capability). I see
+basically three ways to do this, all based on the idea of enforcing that the
+callback should be constructed in a part of memory for which no store-local
+permissions are available:
+\begin{itemize}
+\item One way to exclude the scenario is to require that callbacks are provided
+  as non-local capabilities. The downside of this is that local callbacks can be
+  useful for the caller to prevent the callee from storing them.
+\item Another way to exclude the scenario is to require that the stack is
+  allocated in a fixed part of the address space and to check that callbacks
+  point outside of this region before invoking them.
+\item Perhaps we should require that store-local permissions cannot be removed
+  from a capability and simply require that callback pointers do not have
+  store-local set. Perhaps we can allow store-local permissions to be given up,
+  but only if the corresponding part of memory is fully zeroed in the process
+  (or at least all local capabilities stored in the region).
+\end{itemize}
+
+\subsection{A result to prove...}
+\label{sec:os-less-stack-property}
+
+The simplest thing that comes to mind as a formal result for all of the above is
+to look at a concrete program that clearly relies on properties like
+well-bracketed control flow and encapsulation of local variables and prove it
+correct. As a concrete example: we might show an assembly program that
+corresponds to the following (a higher-order program that crosses trust
+boundaries and relies on local variable encapsulation and well-bracketed control
+flow):
+\begin{verbatim}
+let trustedCode = fun adversary =>
+                    let x = ref 0 in
+                    let callback = fun adv2 => let y = ref (!x) in assert (!x >= 0); x := -1; c2 unit; x := !y)
+                    let _ = adversary callback
+                    assert (!x >= 0)
+\end{verbatim}
 

### 2016-09-22 capability-machine.tex
Added a comment.

@@
 \begin{document}
+\begin{flushright}
+\today
+\end{flushright}
 $\RegName$ contains $\pcreg$, but is otherwise some undefined, finite set.
@@
                     let x = ref 0 in
-                    let callback = fun adv2 => let y = ref (!x) in assert (!x >= 0); x := -1; c2 unit; x := !y)
+                    let callback = fun adv2 => 
+                      let y = ref (!x) in 
+                      assert (!x >= 0); 
+                      x := -1; 
+                      c2 unit; 
+                      x := !y)
                     let _ = adversary callback
@@
 \end{verbatim}
-
+\lau{I have inserted some line breaks for readability. I am not sure what is going on here (the parenthesis after y is unmatched.)}
 \section{Related reading}
@@
 \begin{appendices}
-\section{Definitoins and Proofs related to Section~\ref{subsec:recursive-dom-eq}}
+\section{Definitions and Proofs related to Section~\ref{subsec:recursive-dom-eq}}
 \end{appendices}

### 2016-09-23 capability-machine.tex
Some small fixes and some comments.

@@
 
+\dominique{I suppose you're using a standard ultrametric metric to make an
+  o.f.e. a metric space?}
 \begin{definition}[\cofes]
@@
 We denote this category $\CatC$.
-The category of preordered \cofes{} has preordered \cofes as objects and monotone and non-expansive functions as morphisms.
+The category of preordered \cofes{} has preordered \cofes{} as objects and monotone and non-expansive functions as morphisms.
 We denote this category $\CatP$.
@@
 
-We need the following constructions to create the preordered \cofe{}, we want to work use
+We need the following constructions to create the preordered \cofe{}, we want to
+work use \dominique{this sentence doesn't parse :)}
 \begin{definition}[Preordered \cofe{}\ construction: Function]
@@
 
-The above are standard constructions, so they are used here without showing they infact are well-defined.
+The above are standard constructions, so they are used here without showing they
+are in fact well-defined.
 
@@
 
+\dominique{What is HS? Do you mean $\HeapSegments$?}
+\dominique{Why the exponential in the definition of $G$, why not $W^{\States} \monnefun \UPred{HS}$?}
 Define functors $K$, $R$, and $G$ as follows:
@@
 \end{lemma}
-With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-copmosition of $blater$:
+With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-composition of $\blater$:
 \[
@@
 \]
-The solution to the recursice domain equations is presented \cite{Birkedal:2010:TCS:411:4102-4122}.
+The solution to the recursice domain equations is presented by \cite{Birkedal:2010:TCS:411:4102-4122}.
 They solve it in pre-ordered, non-empty, complete, 1-bounded ultrametric spaces, but they have a simple correspondence to pre-ordered \cofes{}.
@@
 Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
+\dominique{I don't understand the above remark: why should an entry pointer be
+  in the expression relation, since it shouldn't be able to reach the pc?}
 \lau{In the ticket dispenser lemma, we need to make some assumptions about the adversary to be able to use the FTLR.

### 2016-09-23 capability-machine.tex
Fix mistake.

@@
                       x := -1; 
-                      c2 unit; 
+                      adv2 unit; 
                       x := !y)

### 2016-09-23 capability-machine.tex
merge

--- 2016-09-23 capability-machine.tex
@@
 \newcommand{\Worlds}{\plaindom{World}}
+\newcommand{\Wor}{\plaindom{Wor}}
 \newcommand{\StorePred}{\plaindom{HeapSegPred}}
@@
 \[
-W = \nats \finparfun (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments})))
+\Wor = \nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
 \]
@@
 \[
-  \xi : W \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments}))))
+  \xi : \Wor \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
 \]
@@
 \end{lemma}
-With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-composition of $\blater$:
-\[
-  F'(W) = \blater (F(W))
+With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-copmosition of $\blater$:
+\[
+  F'(\Wor) = \blater (F(\Wor))
 \]
@@
 \[
-  W \cong F'(W) = \blater (\nats \finparfun (\States \times \Rel \times (\States \fun W \monnefun \UPred{\HeapSegments} )))
+  \Wor \cong F'(\Wor) = \blater (\nats \finparfun (\States \times \Rel \times (\States \fun \Wor \monnefun \UPred{\HeapSegments} )))
 \]
@@
 \subsection{Worlds}
-Assume preordered \cofe{}\ $W$ and isomorphism $\xi$ such that:
-\[
-  \xi : W \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments}))))
+Assume preordered \cofe{}\ $\Wor$ and isomorphism $\xi$ such that:
+\[
+  \xi : \Wor \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
 \]
@@
 \[
-\Regions \defeq (\States \times \Rel \times (\States \fun (W \monnefun \UPred{\HeapSegments})))
+\Regions \defeq (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
 \]
@@
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) &&\text{ iff } & &
-(s_1,\phi_1) = (s_2,\phi_2) \text{ and } (s_1,s_2) \in \phi_2
+(\phi_1,H_1) = (\phi_2,H_2) \text{ and } (s_1,s_2) \in \phi_2
   \end{align*}
@@
     \text{ and }\\
-    \forall \hat{W} \in \blater \Worlds \ldotp H_1 \; s_1 \; \hat{W} \nsubeq H_2 \; s_2 \; \hat{W}
+    \forall W \in \Wor \ldotp H_1 \; s_1 \; W \nsubeq H_2 \; s_2 \; W
   \end{gathered}
@@
 \subsection{Well-foundedness lemmas}
-\begin{lemma}
+\begin{lemma}[Future worlds transitive]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Value relation uniformity]
 \label{lem:stdvr-uniform-step}
@@
 
-\begin{lemma}
+\begin{lemma}[Value relation monotone in worlds]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Observation relation uniformity]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Register relation uniformity]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Register relation montone in worlds]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Continuation relation uniformity]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Continuation relation monotone in worlds]
   \begin{align*}
@@
 
-\begin{lemma}
+\begin{lemma}[Expression relation uniformity]
 \label{lem:stder-mon-step}

### 2016-09-23 capability-machine.tex
Fix well-bracketedness example.

--- 2016-09-23 capability-machine.tex
@@
                     let callback = fun adv2 => 
+                      x := !x + 1; 
                       let y = ref (!x) in 
-                      assert (!x >= 0); 
-                      x := -1; 
                       adv2 unit; 
-                      x := !y)
+                      assert (!x == !y); 
+                      x := !x - 1)
                     let _ = adversary callback
-                    assert (!x >= 0)
+                    assert (!x == 0)
 \end{verbatim}

### 2016-09-23 capability-machine.tex

--- 2016-09-23 capability-machine.tex
@@
                     let callback = fun adv2 => 
+                      x := !x + 1; 
                       let y = ref (!x) in 
-                      assert (!x >= 0); 
-                      x := -1; 
-                      c2 unit; 
-                      x := !y)
+                      adv2 unit; 
+                      assert (!x == !y); 
+                      x := !x - 1)
                     let _ = adversary callback
-                    assert (!x >= 0)
+                    assert (!x == 0)
 \end{verbatim}

### 2016-10-04 capability-machine.tex
added preservation of limits to preordered cofe def, small clarifications

--- 2016-09-23 capability-machine.tex
@@
 
-\declaretheorem[numbered=yes,name=Lemma]{lemma}
-\declaretheorem[numbered=yes,name=Definition]{definition}
-\declaretheorem[numbered=yes,name=Specification]{specification}
+\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
+\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
+\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}
 
@@
 \newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
+\newcommand\kars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
 
@@
 \begin{definition}[Preordered \cofes]
-  A preordered \cofe{}\ is a \cofe{}\ equiped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. That is the ordering must satisfy the following
+  A preordered \cofe{}\ is a \cofe{}\ equiped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
   \begin{itemize}
-  \item $\forall x \in X \ldotp x \future x$
-  \item $\forall x,y,z \in X \ldotp x \future y \land y \future z \Rightarrow x \future z$
+  \item The ordering preserves limits. That is, for Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
   \end{itemize}
+  % Preorder:
+  % \begin{itemize}
+  % \item $\forall x \in X \ldotp x \future x$
+  % \item $\forall x,y,z \in X \ldotp x \future y \land y \future z \Rightarrow x \future z$
+  % \end{itemize}
 \end{definition}
@@
 
-We need the following constructions to create the preordered \cofe{}, we want to
+We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
 work use \dominique{this sentence doesn't parse :)}
@@
 \end{lemma}
+\begin{proof}
+  %TODO
+\end{proof}
+
+\begin{lemma}
+\label{lem:nequal-and-future-world}
+  \begin{align*}
+    & \forall W_1,W_2,W_1' \ldotp \\
+    & \quad W_1 \nequal W_2 \land W_1' \future W_1 \Rightarrow \\
+    & \qquad \exists W_2' \ldotp W_2' \future W_2 \land W_1' \nequal W_2'
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Assume $W_1 \nequal W_2$ and $W_1' \future W_2'$. Define $W_2'$ to have $\dom(W_2')=\dom(W_1')$ where
+\[
+  W_2' \; r = \left\{
+  \begin{aligned}
+    & (s_1',\phi_2,H_2) & &\text{for } j \in \dom(W_2) \text{ and } W_2 \; j = (s_2,\phi_2,H_2) \\
+    &                  & &\quad \text{ and } W_1' \; j = (s_1',\phi',H_1') \\
+    & W_1' \; r        & &\text{otherwise}
+  \end{aligned} 
+  \right.
+\]
+
+First we show $W_1' \nequal W_2'$.
+The domains are defined to be equal, so that is satsfied.
+Let $r \in \dom(W_1')$ and show $W_1' \; r \nequal W_2' \; r$.
+First consider the case where $r \in \dom(W_2)$.
+Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$.
+We need to show
+\begin{align*}
+  s_1' & \nequal s_1' \\
+  \phi_1' & \nequal \phi_2 \\
+  H_1' & \nequal H_2
+\end{align*}
+The first one is trivial.
+For the second one we use that we know $\phi_1 \nequal \phi_2$ and $\phi_1 = \phi_1'$ respectively from the $n$-equality assumption and future world assumption.
+Similarly, the second one follows from $H_1 \nequal H_2$ and $H_1 = H_1'$ - again respectively from the $n$-equality assumption and the future world assumption.
+For the case where $r \not\in \dom(W_2)$, we need to show $W_1' \; r \nequal W_1' \; r$ which follows trivially from reflexivity.
+
+Next we show $W_2' \future W_2$.
+We know that $\dom(W_2') = \dom(W_1')$, $\dom(W_1') \supseteq \dom(W_1)$ (from future world assumption), and $\dom(W_1) = \dom(W_2)$ (from $n$-equality assumption).
+All this can easily be put together to get $\dom(W_2') \supseteq \dom(W_2)$.
+Now let $r\in \dom(W_2)$ and assume Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$, then we need to show the following:
+\begin{align*}
+  &(s_2,s_1') \in \phi_2 \\
+  &\phi_2 = \phi_2 \\
+  &H_2 = H_2
+\end{align*}
+Here the two last are trivial.
+The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notive that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  %TODO
+\end{proof}
 
@@
 
+\begin{lemma}[Read condition non-expansive in worlds]
+  \begin{align*}
+    & \forall b,e, n, W, W' \ldotp \\
+    & \quad W \nequal W' \Rightarrow \\
+    & \qquad \forall k < n \ldotp \readCond{k,W,b,e}\\
+    & \qquad \quad readCon{k,W',b,e}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  %TODO
+\end{proof}
+
 \begin{lemma}[Read-write condition uniformity]
@@
 
+\begin{lemma}[Execute condition non-expansive in worlds]
+\label{lem:exec-cond-ne}
+  \begin{align*}
+    &\forall W_1, W_2,n,b,e,p \ldotp \\
+    &\quad W_1 \nequal W_2 \Rightarrow \\
+    &\qquad \forall k \leq n \ldotp \\
+    &\qquad \quad \execCond{k,W_1,b,e,p} \\
+    &\qquad \qquad\Rightarrow \execCond{k,W_2,b,e,p}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Assume $W_1 \nequal W_2$. Let $k \leq n$ be given. Assume $\execCond{k,W_1,b,e,p}$. Let $k' < k$, $W_2' \future W_2$, and $a \in [b,e]$ be given. We need to show:
+\[
+  \npair[k']{(p,b,e,a)} \in \stder(W_2')
+\]
+By lemma~\ref{lem:nequal-and-future-world} there exists a future world $W_1'$ such that $W_1' \nequal W_2'$ and $W_1' \future W_1$.
+Now use the execute condition assumption with $k' < k$, $W_1' \future W_1$, and $a$ to get
+\[
+  \npair[k']{(p,b,e,a)} \in \stder(W_1')
+\]
+Be lemma~\ref{lem:stder-ne}, we know $\stder$ is non-expansive, so $\stder(W_1') \nequal \stder(W_2')$. As $k' < n$, we got the result we wanted.
+\end{proof}
+
 \begin{lemma}[Entry condition uniformity]
@@
 
+\begin{lemma}[Entry condition non-expansive in worlds]
+  \begin{align*}
+    &\forall W_1, W_2,n,b,e,a \ldotp \\
+    &\quad W_1 \nequal W_2 \Rightarrow \\
+    &\qquad \forall k \leq n \ldotp \\
+    &\qquad \quad \entryCond{k,W_1,b,e,a} \\
+    &\qquad \qquad\Rightarrow \entryCond{k,W_2,b,e,a}
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  Similar to the proof of Lemma~\ref{lem:exec-cond-ne}.
+\end{proof}
+
 \begin{lemma}[Write condition implies read condition]
@@
 That is Lemma~\ref{lem:readCond-mono-world},~\ref{lem:writeCond-mono-world},~\ref{lem:execCond-mono-world}, and~\ref{lem:entryCond-mono-world}.
+\end{proof}
+
+\begin{lemma}[Value relation contractive]
+  \begin{align*}
+    & \forall n \ldotp \forall W \in \Worlds \forall w \ldotp \\
+    & \quad \npair[n]{w} \in \stdvr(W) \Rightarrow \\
+    & \qquad \npair[n+1]{w} \in \stdvr(W)
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  %TODO
 \end{proof}
@@
 \end{proof}
+
+\begin{lemma}[Expression relation non-expansive in world]
+\label{lem:stder-ne}
+\begin{align*}
+  \forall W_1 \nequal W_2 \ldotp \stder(W_1) \nequal \stder(W_2)
+\end{align*}
+\end{lemma}
 

### 2016-10-18 capability-machine.tex
Updated structure of introduction of the logical relation (still work in progress, so the same thing is repeated), so it was easier to prove well-foundedness and the existense of the value relation.

@@
 \newcommand{\finparfun}{\stackrel{\textit{\tiny{fin}}}{\rightharpoonup}}
-\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\longrightarrow}}
-\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\longrightarrow}}
-\newcommand{\nefun}{\stackrel{\textit{\tiny{ne}}}{\longrightarrow}}
-\newcommand{\fun}{\longrightarrow}
+\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\rightarrow}}
+\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\rightarrow}}
+\newcommand{\nefun}{\stackrel{\textit{\tiny{ne}}}{\rightarrow}}
+\newcommand{\fun}{\rightarrow}
 \newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
@@
 \newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
+\newcommand{\DCPred}[1]{\plaindom{P}^\downarrow(#1)}
 
@@
                            \end{cases} \\
-  \updatePcPerm{(\perm,\start,\addrend,\addr)} &=
+  \updatePcPerm{w} &=
                                      \begin{cases}
-                                       (\exec,\start,\addrend,\addr) & \text{if $\perm = \entry$}\\
-                                       (\perm,\start,\addrend,\addr) & \text{otherwise} 
+                                       (\exec,\start,\addrend,\addr) & \text{if $w = (\entry,\start,\addrend,\addr)$}\\
+                                       w & \text{otherwise} 
                                      \end{cases} \\
@@
 
-The category of \cofes{} is the category whith \cofes{} as objects and non-expansive functions as morphisms.
-We denote this category $\CatC$.
-The category of preordered \cofes{} has preordered \cofes{} as objects and monotone and non-expansive functions as morphisms.
-We denote this category $\CatP$.
-
 \begin{definition}[Preordered \cofe{}\ construction: Finite-partial function]
@@
 We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
-work use \dominique{this sentence doesn't parse :)}
+\dominique{this sentence doesn't parse :)}
 \begin{definition}[Preordered \cofe{}\ construction: Function]
@@
 The above are standard constructions, so they are used here without showing they
-are in fact well-defined.
+are in fact well-defined as shown in \cite{Birkedal:tutorial-notes}.
 
@@
 
-\dominique{What is HS? Do you mean $\HeapSegments$?}
-\dominique{Why the exponential in the definition of $G$, why not $W^{\States} \monnefun \UPred{HS}$?}
+The category of \cofes{} is the category whith \cofes{} as objects and non-expansive functions as morphisms.
+We denote this category $\CatC$.
+The category of preordered \cofes{} has preordered \cofes{} as objects and monotone and non-expansive functions as morphisms.
+We denote this category $\CatP$.
+
 Define functors $K$, $R$, and $G$ as follows:
@@
   & G : \CatP^{\var{op}} \rightarrow \CatC \\
-  & G(W) = \UPred{HS}^{W^{\States}} \\
+  & G(W) = \States \nefun W \monnefun \UPred{HS} \\
   & G(g) = \lambda H \ldotp \lambda \var{st} \ldotp \lambda x \ldotp H (\var{st})(g(x))
@@
 \]
-and we define worlds as
+define region names to be natural numbers, i.e.,
+\[
+  \RegionNames \defeq \nats
+\]
+and define worlds as
 \[
@@
 
+\subsection{Logical Relation}
+\label{subsec:logical-relation}
+Our logical relation is defined using multiple recursive definitions, so the definitions in the following subsections are defined simultaneously.
+We want to define the value relation as the fixed-point given by Banach's fixed-point theorem, so all our definitions will be parameterized with the value relation.
+\subsubsection{Observation Relation}
+\label{subsubsec:observation-relation}
+In order to define the expression relation, we define an observation relation.
+\begin{align*}
+  \observations \quad : & \quad  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
+  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
+                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                           & \qquad \Rightarrow \exists W' \future W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
+\end{align*}
+A pair of a register and a heap segment is ``good'' if we can put it together with a frame heap, so we can execute it.
+The execution should then end up in a heap where the frame remains the same and the remaining heap segment satisfies the world.
+
+Note that the operational semantic is total, so we cannot get stuck.
+If the execution ends up in a $\failed$ configuration, then we do not care about the heap and the registers.
+This is why, we only have requirements on the result when we end up in a $\halted$ configuration.
+
+The following lemmas show that the observation relation is well-defined.
+\begin{lemma}[Observation relation uniformity]
+\label{lem:obs-dc}
+  \begin{align*}
+    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp\\
+    &\quad \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Observation relation non-expansive in worlds]
+\label{lem:obs-ne-worlds}
+\begin{align*}
+  &\forall W, W', n \ldotp \\
+  &\quad W \nequal W' \Rightarrow \observations(W) \nequal \observations(W')
+\end{align*}
+\end{lemma}
+
+\subsubsection{Regiser-File Relation}
+\label{subsubsec:register-file-relation}
+This relation is used in the definition of the continuation relation as well as the expression relation.
+\begin{align*}
+  \stdrr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Regs} \\
+  \stdrr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp
+                     \begin{aligned}[t]
+                       \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                                            & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
+                     \end{aligned}
+\end{align*}
+Well-formedness lemmas for this definition:
+\begin{lemma}[Register relation uniformity]
+\label{lem:reg-dc}
+  \begin{align*}
+    &\forall \stdvr, n' < n \ldotp \forall W \ldotp \forall \reg \ldotp\\
+    &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair[n']{\reg} \in \stdrr(\stdvr)(W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Register relation montone in worlds]
+\label{lem:reg-mono-worlds}
+  \begin{align*}
+    &\forall \stdvr, n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp\\
+    &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair{\reg} \in \stdrr(\stdvr)(W')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Register relation non-expansive in value relation]
+\label{lem:reg-ne-vr}
+\begin{align*}
+  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stdrr(\stdvr) \nequal \stdrr(\stdvr')
+\end{align*}
+\end{lemma}
+
+\subsubsection{Continuation Relation}
+\label{subsubsection:continuation-relation}
+The continuation relation is used in the definition of the expression relation.
+The continuation relation ensures that if you continue execution through a continuation, then it will result in a good result according to the world.
+\begin{align*}
+  \stdkr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
+  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
+                     \begin{aligned}[t]
+                       & \npair{c} \in \stdvr(W) \land \\
+                       &  \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
+                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(\stdvr)(W') \ldotp \\
+                       & \quad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
+                 \end{aligned}
+\end{align*}
+
+Well-definedness lemmas:
+\begin{lemma}[Continuation relation uniformity]
+\label{lem:stdkr-dc}
+  \begin{align*}
+    &\forall \stdvr \ldotp \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp\\
+    &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair[n']{c} \in \stdkr(\stdvr)(W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Continuation relation monotone in worlds]
+\label{lem:stdkr-mono-worlds}
+  \begin{align*}
+    &\forall \stdvr \ldotp \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp\\
+    &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair{c} \in \stdkr(\stdvr)(W')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Continuation relation non-expansive in value relation]
+\label{lem:stdkr-ne-vr}
+\begin{align*}
+  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stdkr(\stdvr) \nequal \stdkr(\stdvr')
+\end{align*}
+\end{lemma}
+
+\subsubsection{Expression Relation}
+\label{subsubsection:expression-relation}
+The expression relation is defined as follows:
+\begin{align*}
+  \stder \quad : & \quad  (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \nefun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
+  \stder \defeq & \lambda \stdvr \ldotp \lambda W \ldotp \{ \npair{\pc} \mid 
+  \begin{aligned}[t]
+    & \forall n' \leq n \ldotp\\
+    & \quad \forall \npair[n']{\reg} \in \stdrr(\stdvr)(W) \ldotp \\
+    & \quad \quad \forall \npair[n']{c} \in \stdkr(\stdvr)(W) \ldotp \\
+    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
+    & \quad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
+  \end{aligned}
+\end{align*}
+Well-definedness lemmas:
+\begin{lemma}[Expression relation uniformity]
+\label{lem:stder-dc}
+  \begin{align*}
+    & \forall \stdvr \ldotp \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp\\
+    & \quad \npair{\pc} \in \stder(\stdvr)(W) \Rightarrow \npair[n']{\pc} \in \stder(\stdvr)(W) 
+ \end{align*}
+\end{lemma}
+\begin{lemma}[Expression relation non-expansive in world]
+\label{lem:stder-ne-worlds}
+\begin{align*}
+  \forall \stdvr \ldotp \forall W_1 \nequal W_2 \ldotp \stder(\stdvr)(W_1) \nequal \stder(\stdvr)(W_2)
+\end{align*}
+\end{lemma}
+
+\begin{lemma}[Expression relation non-expansive in value relation]
+\label{lem:stder-ne-vr}
+\begin{align*}
+  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stder(\stdvr) \nequal \stder(\stdvr')
+\end{align*}
+\end{lemma}
+
+\subsubsection{Standard Region}
+\label{subsubsec:standard-region}
+The following standard region is used in the definition of the value relation. Specifically, it is used in the $\readCond{}$ and the $\writeCond{}$ (to be defined next)
+
+\begin{align*}
+  \iota_{\var{start},\var{end}} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun \Regions\\
+  \iota_{\var{base},\var{end}} & \defeq \lambda \; \stdvr \ldotp ((\var{base},\var{end}),=,H_{\var{std}} (\stdvr)) \\ & \\
+  H_{\var{std}} & : (\Worlds \monnefun \UPred{\Words}) \nefun \States \nefun \Worlds \monnefun \UPred{\HeapSegments} \\
+  H_{\var{std}} \; \stdvr \; (\var{base},\var{end})\; W &  \defeq \left\{ \npair{\hs} \middle|
+                                               \begin{aligned}
+                                                 & \dom(\hs) = [\var{base},\var{end}] \land \\
+                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; W)
+                                               \end{aligned} \right\}
+\end{align*}
+As mentioned previously, the set of states contains the ``necessary'' states. For the above to make sense, the set of states contains pairs of natural numbers $(\var{base},\var{end})$.
+
+The well-definedness lemmas for the above is:
+\begin{lemma}[$H_{\var{std}}$ is monotone in the worlds]
+\label{lem:std-reg-mono-world}
+\begin{align*}
+  & \forall \stdvr \ldotp \forall \start, \addrend \ldotp \forall W' \future W \ldotp \\
+  & \quad H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W' \supseteq H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W
+\end{align*}
+\end{lemma}
+
+\begin{lemma}[$H_{\var{std}}$ is non-expansive in the worlds]
+\label{lem:std-reg-ne-world}
+\begin{align*}
+  & \forall \stdvr \ldotp \forall \start, \addrend \ldotp \forall n \ldotp \forall W_1 \nequal W_2 \ldotp \\
+  & \quad H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W_1 \nequal H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W
+\end{align*}
+\end{lemma}
+
+\begin{lemma}[$H_{\var{std}}$ is non-expansive in the value relation]
+\label{lem:std-reg-ne-vr}
+\begin{align*}
+  & \forall \stdvr, \stdvr' \ldotp \forall n \ldotp \stdvr \nequal \stdvr' \Rightarrow\\
+  & \quad H_{\var{std}}\; \stdvr \nequal H_{\var{std}}\; \stdvr'
+\end{align*}
+\end{lemma}
+
+\begin{lemma}[$\iota_{\start,\addrend}$ is non-expansive in the value relation]
+\label{lem:std-reg-i-ne-vr}
+\begin{align*}
+  & \forall \start, \addrend \ldotp \forall \stdvr, \stdvr' \ldotp \forall n \ldotp \stdvr \nequal \stdvr' \Rightarrow\\
+  & \quad \iota_{\start,\addrend}\; \stdvr \nequal \iota_{\start,\addrend}\; \stdvr'
+\end{align*}
+\end{lemma}
+
+\subsubsection{Capability Conditions}
+\label{subsubsec:capability-conditions}
+The definition of the value relation has the same conditions several times, so to define it consisely, we define the following conditions.
+\begin{align*}
+  \readCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
+  \readCond{}(\stdvr)(\start,\addrend,W) & =        
+  \begin{aligned}[t]
+    \{ n \mid & \;\exists r \in \RegionNames \ldotp \\
+              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+              & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'}(\stdvr) \}
+  \end{aligned}\\
+ & \\
+  \writeCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
+  \writeCond{}(\stdvr)(\start,\addrend, W) & = 
+  \begin{aligned}[t]
+    \{ n \mid & \; \exists r \in \RegionNames \ldotp \\
+              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+              & \; \qquad W(r) \nequal[n-1] \iota_{\start',\addrend'}(\stdvr)  \}
+  \end{aligned}\\
+ & \\
+  \execCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Perms \times \Worlds) \monnefun \DCPred{\nats} \\
+  \execCond{}(\stdvr)(\start,\addrend,\perm,W) & = 
+  \begin{aligned}[t]
+    \{ n \mid & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
+              & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
+              & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W') \}
+  \end{aligned} \\
+ & \\
+  \entryCond{} & : \;  (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^3 \times \Worlds) \monnefun \DCPred{\nats} \\
+  \entryCond{}(\stdvr)(\start,\addrend,\addr,W) & =
+  \begin{aligned}[t]
+    \{ n \mid & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
+              & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(\stdvr)(W') \}
+  \end{aligned}
+\end{align*}
+
+The following lemmas show that the above conditions are well-defined:
+\begin{lemma}[Read condition downwards-closed]
+\label{lem:rc-dc}
+  \begin{align*}
+    &\forall \stdvr, n,n',W,\start, \addrend \ldotp \\
+    &\quad  n \in \readCond{}(\stdvr)(\start,\addrend,W) \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow n' \in \readCond{}(\stdvr)(\start,\addrend,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Read condition monotone in world]
+\label{lem:rc-mono-world}
+  \begin{align*}
+    &\forall \stdvr, n,W,W',\start, \addrend \ldotp \\
+    &\quad  (\start,\addrend,W') \future (\start,\addrend,W) \\
+    &\qquad \Rightarrow \readCond{}(\stdvr)(\start,\addrend,W') \supseteq \readCond{}(\stdvr)(\start,\addrend,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Read condition non-expansive in worlds]
+\label{lem:rc-ne-world}
+  \begin{align*}
+    & \forall \stdvr, b,e, n, W, W' \ldotp \\
+    & \quad (b,e,W) \nequal (b,e,W') \Rightarrow \\
+    & \qquad \readCond{}(\stdvr)(b,e,W) \nequal \readCond{}(\stdvr)(b,e,W')\\
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Read-write condition uniformity]
+\label{lem:wc-dc}
+  \begin{align*}
+    &\forall \stdvr, n,n',W,\start, \addrend \ldotp \\
+    &\quad  n \in \writeCond{}(\stdvr)(\start,\addrend,W) \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow n' \in \writeCond{}(\stdvr)(\start,\addrend,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Read-write condition monotone in world]
+\label{lem:wc-mono-world}
+  \begin{align*}
+    &\forall \stdvr, n,W,W',\start, \addrend \ldotp \\
+    &\quad  (\start,\addrend,W') \future (\start,\addrend,W) \\
+    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start,\addrend,W') \supseteq \writeCond{}(\stdvr)(\start,\addrend,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Execute condition downwards-closed]
+\label{lem:ec-dc}
+  \begin{align*}
+    &\forall \stdvr, n,n',W,\start, \addrend, \perm \ldotp \\
+    &\quad  n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W) \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow n' \in \execCond{}(\stdvr)(\start,\addrend,\perm,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Execute condition monotone in world]
+\label{lem:ec-mono-world}
+  \begin{align*}
+    &\forall \stdvr,n,W, W',\start, \addrend, \perm \ldotp \\
+    &\quad  (\start,\addrend,\perm,W') \future (\start,\addrend,\perm,W) \\
+    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\perm,W') \supseteq \execCond{}(\stdvr)(\start,\addrend,\perm,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Execute condition non-expansive in worlds]
+\label{lem:ec-ne-world}
+  \begin{align*}
+    &\forall \stdvr, W_1, W_2,n,\start,\addrend,\perm \ldotp \\
+    &\quad (\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2) \Rightarrow \\
+    &\qquad \Rightarrow \execCond{}(\stdvr)(b,e,p,W_1) \nequal \execCond{}(\stdvr)(b,e,p,W_2)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Entry condition downwards-closed]
+\label{lem:en-dc}
+  \begin{align*}
+    &\forall \stdvr, n,n',W,\start, \addrend, \addr \ldotp \\
+    &\quad  n \in \entryCond{}(\stdvr)(\start,\addrend,\addr,W) \land \\
+    &\quad  n' \leq n \\
+    &\qquad \Rightarrow n' \in \entryCond{}(\stdvr)(\start,\addrend,\addr,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Entry condition monotone in world]
+  \label{lem:en-mono-world}
+  \begin{align*}
+    &\forall \stdvr,n,W, W',\start, \addrend, \addr \ldotp \\
+    &\quad  (\start,\addrend,\addr,W') \future (\start,\addrend,\addr,W) \\
+    &\qquad \Rightarrow \entryCond{}(\stdvr)(\start,\addrend,\perm,W') \supseteq \entryCond{}(\stdvr)(\start,\addrend,\perm,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Entry condition non-expansive in worlds]
+\label{lem:en-ne-world}
+  \begin{align*}
+    &\forall \stdvr, W_1, W_2,n,\start,\addrend,\addr \ldotp \\
+    &\quad (\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2) \Rightarrow \\
+    &\qquad \Rightarrow \execCond{}(\stdvr)(b,e,p,W_1) \nequal \execCond{}(\stdvr)(b,e,p,W_2)
+  \end{align*}
+\end{lemma}
+Finally, we need to show that all the conditions are non-expansive, but we later want to use Banach's fixed point theorem to define the value relation.
+For this we will need that the above conditions are contractive, and if they are contractive, then they are also non-expansive, so we show that each of the conditions are contractive:
+\begin{lemma}[Read condition contractive]
+  \label{lem:rc-con}
+  \begin{align*}
+    &\forall \stdvr, \stdvr', n \ldotp \\
+    &\quad \stdvr \nequal \stdvr' \Rightarrow \readCond{}(\stdvr) \nequal[n+1] \readCond{}(\stdvr')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Write condition contractive]
+  \label{lem:wc-con}
+  \begin{align*}
+    &\forall \stdvr, \stdvr', n \ldotp \\
+    &\quad \stdvr \nequal \stdvr' \Rightarrow \writeCond{}(\stdvr) \nequal[n+1] \writeCond{}(\stdvr')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Execute condition contractive]
+  \label{lem:ec-con}
+  \begin{align*}
+    &\forall \stdvr, \stdvr', n \ldotp \\
+    &\quad \stdvr \nequal \stdvr' \Rightarrow \execCond{}(\stdvr) \nequal[n+1] \execCond{}(\stdvr')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Entry condition contractive]
+  \label{lem:en-con}
+  \begin{align*}
+    &\forall \stdvr, \stdvr', n \ldotp \\
+    &\quad \stdvr \nequal \stdvr' \Rightarrow \entryCond{}(\stdvr) \nequal[n+1] \entryCond{}(\stdvr')
+  \end{align*}
+\end{lemma}
+
+\subsubsection{Value Relation}
+\label{subsubsec:value-relation}
+The value relation, is defined as follows:
+\begin{align*}
+  \stdvr & :  (\Worlds \monnefun \UPred{Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
+  \stdvr &\defeq \lambda \; V \ldotp \lambda \; W \ldotp 
+              \begin{aligned}[t]
+  & \{ \npair{i} \mid i \in \ints \} 
+    \union \\
+         & \{ \npair{\stdcap[\noperm] }  \} 
+           \union \\
+         & \{ \npair{\stdcap[\readonly] } \mid n \in \readCond{}(V)(\start,\addrend,W) \} \union \\
+         & \{ \npair{\stdcap[\readwrite] } \mid n \in \writeCond{}(V)(\start,\addrend,W) \} \union \\
+         & \{ \npair{\stdcap[\exec]} \mid \\
+         & \quad n \in \readCond{}(V)(\start,\addrend,W) \land \\
+         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \}
+           \union \\
+           % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
+         & \{ \npair{\stdcap[\entry]} \mid n \in \entryCond{}(V)(\start,\addrend,\addr,W) \} \union \\
+         & \{ \npair{\stdcap[\rwx]} \mid \\
+         & \quad n \in \writeCond{}(V)(\start,\addrend,W) \land \\
+         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \land \\
+         & \quad n \in \execCond{}(V)(\start,\addrend,\rwx,W) \}
+       \end{aligned}
+\end{align*}
+
 \subsection{Standard Regions}
 \label{subsec:standard-regions}
-A standard heap invariant that insures all values in the region are in the value  relation:
+To define the value relation, we use a standard heap invariant that ensures all values in the region are in the value relation.
+The following region uses pairs of natural numbers, $(\var{base},\var{end})$, as states, so pairs of natural numbers are in the set $\States$.
 \[
@@
 \end{align*}
+Note that this region is defined in terms of the value relation, and the value relation is defined in terms of this invariant. We define the well-definedness lemma here, but show it in the appendix.
+
 \begin{lemma}[$\iota_{\var{start},\var{end}}$ is well-defined]
@@
 \subsection{Value Relation}
-In order to define the value relation consiely using the standard regions from the Subsection~\ref{subsec:standard-regions}, we define the following conditions:
+In order to define the value relation concisely using the standard regions from Subsection~\ref{subsec:standard-regions}, we define the following conditions:
 \begin{align*}
@@
 
-\[
-\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
-\]
-\begin{align*}
-  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
+\begin{align*}
+  \stdvr & :  \Worlds \monfun \UPred{\Words} \\
+  \stdvr &\defeq \lambda \; W. 
+              \begin{aligned}[t]
+  & \{ \npair{i} \mid i \in \ints \} 
     \union \\
@@
          & \quad \execCond{n,W,\start,\addrend,\rwx} \}
+       \end{aligned}
 \end{align*}
@@
 \end{align*}
-\lau{Added $n' \leq n$ to be able to prove Lemma~\ref{lem:stder-mon-step}}
+\lau{Added $n' \leq n$ to be able to prove Lemma~\ref{lem:stder-dc}}
 
@@
 
-\begin{lemma}[Read condition uniformity]
-  \begin{align*}
-    &\forall n,n',W,\start, \addrend \ldotp \\
-    &\quad  \readCond{n,W,\start,\addrend} \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow \readCond{n',W,\start,\addrend}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Follows directly from definition: Let $n$, $n'$, $\start$, and  $\addrend$ be given.
-Assume $\readCond{n,W,\start,\addrend}$ and $n' \leq n$.
+\begin{proof}[Proof of lemma~\ref{lem:rc-dc}]
+  Follows directly from definition: Let $\stdvr$, $n$, $n'$, $\start$, and  $\addrend$ be given.
+Assume $n \in \readCond{}(\stdvr)(\start,\addrend,W)$ and $n' \leq n$.
 If $n=n'$, then the result follows trivially, as it is exactly our assumption.
 If $n' < n$, then by assumption we have a region name $r$, interval $[\start',\addrend'] \supseteq [\start,\addrend]$, and the assumption $W(r) \nsubeq[n-1] \iota_{\start',\addrend'}$  given by the read condition assumption.
-Picking $r$ and $[\start',\addrend']$, we need to show  $W(r) \nsubeq[n'-1] \iota_{\start',\addrend'}$ which follows from downwards closure of $\nsubeq$.
-\end{proof}
-
-\begin{lemma}[Read condition monotone in world]
-\label{lem:readCond-mono-world}
-  \begin{align*}
-    &\forall n,W,W',\start, \addrend \ldotp \\
-    &\quad  \readCond{n,W,\start,\addrend} \land \\
-    &\quad  W' \future W \\
-    &\qquad \Rightarrow \readCond{n,W',\start,\addrend}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-Let $n$,$W$, $W'$, $b$, and $e$ be given where $W' \future W$.
-Assume $\readCond{n,W,\start,\addrend}$, this give a $r$ and $[b',e'] \supseteq [b,e]$ such that $W(r) \nsubeq \iota_{b',e'} (r) = ((b',e'),=,H_{\var{std}})$.
+Picking $r$ and $[\start',\addrend']$, we need to show  $W(r) \nsubeq[n'-1] \iota_{\start',\addrend'}$ which follows from downwards closure of $\nsubeq[n-1]$.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:rc-mono-world}]
+Let $\stdvr$, $n$,$W$, $W'$, $b$, and $e$ be given where $W' \future W$.
+Let $n \in \readCond{}(\stdvr)(\start,\addrend,W)$, this gives an $r$ and $[b',e'] \supseteq [b,e]$ such that $W(r) \nsubeq \iota_{b',e'} (r) = ((b',e'),=,H_{\var{std}})$.
 This gives
 \begin{align*}
-  &W(r).\phi = = \\
+  &W(r).\phi \equiv = \\
   &W(r).s    = (b',e') \\
@@
 \begin{align*}
-  & W'(r).\phi = = \\
+  & W'(r).\phi \equiv = \\
   & W'(r) = (b',e') \\
@@
 
-\begin{lemma}[Read condition non-expansive in worlds]
-  \begin{align*}
-    & \forall b,e, n, W, W' \ldotp \\
-    & \quad W \nequal W' \Rightarrow \\
-    & \qquad \forall k < n \ldotp \readCond{k,W,b,e}\\
-    & \qquad \quad readCon{k,W',b,e}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  %TODO
-\end{proof}
-
-\begin{lemma}[Read-write condition uniformity]
-  \begin{align*}
-    & \forall n,n',W,\start, \addrend \ldotp\\
-    &\quad \writeCond{n,W,\start,\addrend} \land\\
-    &\quad  n' \leq n\\
-    &\qquad \Rightarrow \writeCond{n',W,\start,\addrend}
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:rc-ne-world}]
+Assume 
+\[
+  (b,e,W) \nequal (b,e,W')
+\]
+which implies $W \nequal W'$. Now let $(n,\hs) \in \readCond{}(\stdvr)(b,e,W)$ be given. This gives us $r$ and $[b',e'] \supseteq [b,e]$ such that 
+\[
+  W(r) \nsubeq[n-1] \iota_{b',e'}(\stdvr)
+\]
+Using $r$ and $[b',e']$, we need to show
+\[
+  W'(r) \nsubeq[n-1] \iota_{b',e'}(\stdvr)
+\]
+Let $W(r) = (s,\phi,H)$, $W(r) = (s',\phi',H')$, and $\iota_{b',e'} = (s_\iota,\phi_\iota,H_\iota)$.
+From the $n$-equality of our worlds, we get $W(r) \nequal W'(r)$. From this we conclude
+\[
+  (s',\phi') = (s,\phi) = (s_\iota,\phi_\iota)
+\]
+It remains to show
+\[
+  \forall \hat{W} \in \Wor \ldotp H' \; s \; W \nequal[n-1] H_\iota \; s \; W
+\]
+Given $\hat{W}$ we know from $n$-equality of the worlds
+\[
+   H' \; s \; \hat{W} \nequal H \; s \; \hat{W}
+\]
+Given $(k,\hs) \in H' \; s \; W$ where $k < n$ and $\hs$ is a heap segment, by above we have $(k,\hs) \in H \; s \; W$ and from the $\nsubeq[n-1]$ assumption, we get $(k,\hs) \in H_\iota \; s_\iota \; W$.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:wc-dc}]
   Follows directly from the definition.
@@
 
-\begin{lemma}[Read-write condition monotone in world]
-\label{lem:writeCond-mono-world}
-  \begin{align*}
-    & \forall n,W, W',\start, \addrend \ldotp\\
-    &\quad \writeCond{n,W,\start,\addrend} \land\\
-    &\quad  W' \leq W\\
-    &\qquad \Rightarrow \writeCond{n,W',\start,\addrend}
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:wc-mono-world}]
   Similar to the proof of the monotonicity of the read condition.
@@
 
-\begin{lemma}[Execute condition uniformity]
-  \begin{align*}
-    &\forall n,n',W,\start, \addrend, \perm \ldotp \\
-    &\quad  \execCond{n,W,\start,\addrend,\perm} \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow \execCond{n',W,\start,\addrend,\perm}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Follows directly from definition: Let $n$, $n'$, $W$, $\start$, and $\addrend$ be given.
-Assume $\execCond{n,W,\start,\addrend,\perm}$ and $n' \leq n$.
+\begin{proof}[Proof of lemma~\ref{lem:ec-dc}]
+  Follows directly from definition: Let $\stdvr$, $n$, $n'$, $W$, $\start$, and $\addrend$ be given.
+Assume $n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W)$ and $n' \leq n$.
 If $n' = n$, then the result follows directly from assumption.
@@
 
-\begin{lemma}[Execute condition monotone in world]
-\label{lem:execCond-mono-world}
-  \begin{align*}
-    &\forall n,W, W',\start, \addrend, \perm \ldotp \\
-    &\quad  \execCond{n,W,\start,\addrend,\perm} \land \\
-    &\quad  W' \leq W \\
-    &\qquad \Rightarrow \execCond{n,W',\start,\addrend,\perm}
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:ec-mono-world}]
   Follows directly from definition.
@@
 
-\begin{lemma}[Execute condition non-expansive in worlds]
-\label{lem:exec-cond-ne}
-  \begin{align*}
-    &\forall W_1, W_2,n,b,e,p \ldotp \\
-    &\quad W_1 \nequal W_2 \Rightarrow \\
-    &\qquad \forall k \leq n \ldotp \\
-    &\qquad \quad \execCond{k,W_1,b,e,p} \\
-    &\qquad \qquad\Rightarrow \execCond{k,W_2,b,e,p}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Assume $W_1 \nequal W_2$. Let $k \leq n$ be given. Assume $\execCond{k,W_1,b,e,p}$. Let $k' < k$, $W_2' \future W_2$, and $a \in [b,e]$ be given. We need to show:
-\[
-  \npair[k']{(p,b,e,a)} \in \stder(W_2')
-\]
-By lemma~\ref{lem:nequal-and-future-world} there exists a future world $W_1'$ such that $W_1' \nequal W_2'$ and $W_1' \future W_1$.
-Now use the execute condition assumption with $k' < k$, $W_1' \future W_1$, and $a$ to get
-\[
-  \npair[k']{(p,b,e,a)} \in \stder(W_1')
-\]
-Be lemma~\ref{lem:stder-ne}, we know $\stder$ is non-expansive, so $\stder(W_1') \nequal \stder(W_2')$. As $k' < n$, we got the result we wanted.
-\end{proof}
-
-\begin{lemma}[Entry condition uniformity]
-  \begin{align*}
-    & \forall n,n',W,\start, \addrend, \addr \ldotp\\
-    &\quad  \entryCond{n,W,\start,\addrend,\addr} \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow \entryCond{n',W,\start,\addrend,\addr}
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:ec-ne-world}]
+  Assume $(\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2)$ and thus $W_1 \nequal W_2$. Given $n'<n$, $W_1' \future W_1$, and $\addr$, show
+\[
+  \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W_1')
+\]
+by lemma~\ref{lem:nequal-and-future-world} there must exist $W_2'$ such that $W_2' \future W_2$ and $W_2' \nequal W_1'$. By assumption $n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W_2)$ we have
+\[
+\npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W_2')
+\]
+and by non-expansiveness of $\stder(\stdvr)$ (lemma~\ref{lem:stder-ne-worlds}), we get the desired result.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:en-dc}]
   Like the $\execCond{}$ uniformity proof.
@@
 
-\begin{lemma}[Entry condition monotone in world]
-  \label{lem:entryCond-mono-world}
-  \begin{align*}
-    & \forall n,W,W',\start, \addrend, \addr \ldotp\\
-    &\quad  \entryCond{n,W,\start,\addrend,\addr} \land \\
-    &\quad  W' \leq W \\
-    &\qquad \Rightarrow \entryCond{n,W',\start,\addrend,\addr}
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:en-mono-world}]
   Follows directly from definition.
@@
 
-\begin{lemma}[Entry condition non-expansive in worlds]
-  \begin{align*}
-    &\forall W_1, W_2,n,b,e,a \ldotp \\
-    &\quad W_1 \nequal W_2 \Rightarrow \\
-    &\qquad \forall k \leq n \ldotp \\
-    &\qquad \quad \entryCond{k,W_1,b,e,a} \\
-    &\qquad \qquad\Rightarrow \entryCond{k,W_2,b,e,a}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Similar to the proof of Lemma~\ref{lem:exec-cond-ne}.
+\begin{proof}[Proof of lemma~\ref{lem:en-ne-world}]
+  This proof goes like the proof of $\execCond{}$ being non-expansive in the second argument (lemma~\ref{lem:ec-ne-world}).
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:rc-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:wc-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:ec-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:en-con}]
+  %TODO
 \end{proof}
@@
 \begin{lemma}[Write condition implies read condition]
+\label{lem:wc-implies-rc}
   \begin{align*}
@@
   \end{align*}
-\end{lemma}
+\end{lemma}[Proof of lemma~\ref{lem:wc-implies-rc}]
 \begin{proof}
@@
   Follows from uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$ in the worlds.
-That is Lemma~\ref{lem:readCond-mono-world},~\ref{lem:writeCond-mono-world},~\ref{lem:execCond-mono-world}, and~\ref{lem:entryCond-mono-world}.
+That is Lemma~\ref{lem:rc-mono-world},~\ref{lem:wc-mono-world},~\ref{lem:ec-mono-world}, and~\ref{lem:en-mono-world}.
 \end{proof}
@@
 
-\begin{lemma}[Observation relation uniformity]
-  \begin{align*}
-    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp\\
-    &\quad \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:obs-dc}]
   Let $n' < n$, $W$, $\reg$, and $\hs$ be given.
@@
 
-\begin{lemma}[Register relation uniformity]
-  \begin{align*}
-    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp\\
-    &\quad \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Follows by uniformity of the value relation, Lemma~\ref{lem:stdvr-uniform-step}.
-\end{proof}
-
-\begin{lemma}[Register relation montone in worlds]
-  \begin{align*}
-    &\forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp\\
-    &\quad \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:obs-ne-worlds}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:reg-dc}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:reg-mono-worlds}]
   This follows by monotonicity of $\stdvr$ in the worlds.
@@
 
-\begin{lemma}[Continuation relation uniformity]
-  \begin{align*}
-    &\forall n' < n \ldotp \forall W \ldotp \forall c \ldotp\\
-    &\quad \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:reg-ne-vr}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdkr-dc}]
 Follows by using the assumption and transitivity of inequality.
@@
 
-\begin{lemma}[Continuation relation monotone in worlds]
-  \begin{align*}
-    &\forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp\\
-    &\quad \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
-  \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:stdkr-mono-worlds}]
 We need to show two things: The continuation is in the value predicate in the future world and that continuing execution using the continuation is in the observation predicate.
@@
 
-\begin{lemma}[Expression relation uniformity]
-\label{lem:stder-mon-step}
-  \begin{align*}
-    & \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp\\
-    & \quad \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W) 
- \end{align*}
-\end{lemma}
-\begin{proof}
+\begin{proof}[Proof of lemma~\ref{lem:stdkr-ne-vr}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stder-dc}]
   Follows from the assumption.
@@
 
-\begin{lemma}[Expression relation non-expansive in world]
-\label{lem:stder-ne}
-\begin{align*}
-  \forall W_1 \nequal W_2 \ldotp \stder(W_1) \nequal \stder(W_2)
-\end{align*}
-\end{lemma}
-
+\begin{proof}[Proof of lemma~\ref{lem:stder-ne-worlds}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stder-ne-vr}]
+    %TODO
+\end{proof}
 \section{Other examples and applications}
@@
 \section{Proof of Lemmas in Section~\ref{subsec:standard-regions}}
-\begin{proof}[Proof of Lemma~\ref{lem:std-region-wd}]
+\begin{proof}[Proof of Lemma~\ref{lem:std-reg-mono-world}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:std-reg-ne-world}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:std-reg-ne-vr}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of Lemma~\ref{lem:std-reg-i-ne-vr}]
   %TODO

### 2016-10-19 capability-machine.tex
Moved text from only LR section to the new one.

@@
 
-\subsection{Logical Relation}
-\label{subsec:logical-relation}
+\subsection{Logical Relation Setup}
+\label{subsec:logical-relation-setup}
 Our logical relation is defined using multiple recursive definitions, so the definitions in the following subsections are defined simultaneously.
@@
 \end{align*}
+We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.
+
+Normally, it would just be required for the return value to be in the value relation, so the continuation is invoked with this value.
+Here everything that is in the register file is available when the continuation is invoked, so we do not special case on $r_1$ (which is the register that in the calling convention we use should contain the return value).
 
@@
 \end{lemma}
-Finally, we need to show that all the conditions are non-expansive, but we later want to use Banach's fixed point theorem to define the value relation.
+Finally, we need to show that all the conditions are non-expansive, but we later want to use Banach's fixed-point theorem to define the value relation.
 For this we will need that the above conditions are contractive, and if they are contractive, then they are also non-expansive, so we show that each of the conditions are contractive:
@@
 \end{align*}
-
-\subsection{Standard Regions}
-\label{subsec:standard-regions}
-To define the value relation, we use a standard heap invariant that ensures all values in the region are in the value relation.
-The following region uses pairs of natural numbers, $(\var{base},\var{end})$, as states, so pairs of natural numbers are in the set $\States$.
-\[
-  \iota_{\var{start},\var{end}} \; : \; \Regions
-\]
-\begin{align*}
-  \iota_{\var{base},\var{end}} & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
-  H_{\var{std}}\; (\var{base},\var{end})\; W &  \defeq \left\{ \npair{\hs} \middle|
+Using the Banach fixed-point theorem, we can find a fixed point for $\stdvr$ We will use this fixed point in the next section to define the logical relation we use in practice.
+
+\subsection{Logical Relation Definitions}
+\label{subsec:logical-relation-definitions}
+\lau{This section introduces the definitions, we started working with, so it is here so we don't have to rewrite everything.}
+Here we state the definitions, we will use for pratical purposes.
+Assume $V$ is the fixed point given by Banach's fixed point theorem as described in Section~\ref{subsubsec:value-relation}.
+In general the definitions here will be the same as in Section~\ref{subsec:logical-relation-setup}, but but applied to $V$ when applicable.
+As they are essentially the same, we will reuse the names.
+
+\begin{align*}
+  \stdrr & \defeq \stdrr \; V \\
+  \stdkr & \defeq \stdkr \; V \\
+  \stder & \defeq \stder \; V \\
+  \stdvr & \defeq \stdvr \; V 
+\end{align*}
+
+\begin{align*}
+  \iota_{\start,\addrend} & \defeq \iota_{\start,\addrend} \; V \\
+  H_{\var{std}}          & \defeq H_{\var{std}} \; V
+\end{align*}
+
+For natural number $n$, adresses $\start$, $\addrend$, and $\addr$, permission $\perm$, and world $W$, we define the capability conditions as follows
+\begin{align*}
+  \readCond{n,W,\start,\addrend}        & \text{ iff } n \in \readCond{}(V)(\start,\addrend,W)\\
+  \writeCond{n,W,\start,\addrend}       & \text{ iff } n \in \writeCond{}(V)(\start,\addrend,W)\\
+  \execCond{n,W,\start,\addrend,\perm}  & \text{ iff } n \in \execCond{}(V)(\start,\addrend,p,W)\\
+  \entryCond{n,W,\start,\addrend,\addr} & \text{ iff } n \in \entryCond{}(V)(\start,\addrend,\addr,W)
+\end{align*}
+The well-formedness lemmas of course still hold for this new presentation.
+
+\subsection{Further Standard Regions}
+\label{subsec:further-standard-regions}
+We have already defined the standard region $\iota_{\start,\addrend}$ which ensures that all the words in the region are in the value relation.
+See Section~\ref{subsubsec:standard-region}.
+We here define another usefull region, namely the one that requires all the words in the region to be the constant 0.
+\[
+  \iota_{\start,\addrend}^0 \; : \; \Regions
+\]
+\begin{align*}
+  \iota_{\start,\addrend}^0 & \defeq ((\start,\addrend),=,H_{\var{std}}^0) \\
+  H_{\var{std}}^0\; (\start,\addrend)\; W     & \defeq\left\{ \npair{\heapseg} \middle| 
                                                      \begin{aligned}
-                                                 & \dom(\hs) = [\var{base},\var{end}] \land \\
-                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; W)
-                                               \end{aligned} \right\}
-\end{align*}
-Note that this region is defined in terms of the value relation, and the value relation is defined in terms of this invariant. We define the well-definedness lemma here, but show it in the appendix.
-
-\begin{lemma}[$\iota_{\var{start},\var{end}}$ is well-defined]
-\label{lem:std-region-wd}
-  For all $\var{base}$ and $\var{end}$, $H_{\var{std}}\; (\var{base}, \var{end})$ is monotone and non-expansive. $=$ is a reflexive and transitive relation.
-\end{lemma}
-
-Another standard heap invariant that ensures all the words in the region are 0.
-\[
-  \iota_{\var{start},\var{end}}^0 \; : \; \Regions
-\]
-\begin{align*}
-  \iota_{\var{base},\var{end}}^0 & \defeq ((\var{base},\var{end}),=,H_{\var{std}}^0) \\
-  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\left\{ \npair{\heapseg} \middle| 
-                                                     \begin{aligned}
-                                                       & \dom(\heapseg) = [\var{base},\var{end}] \land \\
+                                                       & \dom(\heapseg) = [\start,\addrend] \land \\
                                                        & \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  
@@
 
-\begin{lemma}[$\iota_{\var{base},\var{end}}^0$ is well-defined]
+\begin{lemma}[$\iota_{\start,\addrend}^0$ is well-defined]\forcenewline
 \label{lem:std-z-region-wd}
-    For all $\var{base}$ and $\var{end}$, $H_{\var{std}}^0 \; (\var{base},\var{end})$ is monotone and non-expansive. $=$ is a reflexive and transitive relation.
+\begin{itemize}
+  \item $H_{\var{std}}^0$ is non-expansive.
+  \item For all adresses $\start$
+    and $\addrend$, $ H_{\var{std}}^0 \; (\start,\addrend)$ is monotone and non-expansive.
+  \item  $=$ is a reflexive and transitive relation.
+\end{itemize}
 \end{lemma}
@@
 
-\section{Proof of Lemmas in Section~\ref{subsec:standard-regions}}
+\section{Proof of Lemmas in Section~\ref{subsec:further-standard-regions}}
 \begin{proof}[Proof of Lemma~\ref{lem:std-reg-mono-world}]

### 2016-10-19 capability-machine.tex
Removed the old Logical relations section

@@
 
-\begin{lemma}[$\iota_{\start,\addrend}^0$ is well-defined]\forcenewline
-\label{lem:std-z-region-wd}
+\begin{lemma}[$\iota_{\start,\addrend}^0$ is well-defined]\label{lem:std-z-region-wd}\forcenewline
 \begin{itemize}
@@
 
-\subsection{Value Relation}
-In order to define the value relation concisely using the standard regions from Subsection~\ref{subsec:standard-regions}, we define the following conditions:
-\begin{align*}
-  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionNames \ldotp \\
-                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp\\
-                                          & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
-  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionNames \ldotp \\
-                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
-  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
-                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
-  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
-\end{align*}
-
-\begin{align*}
-  \stdvr & :  \Worlds \monfun \UPred{\Words} \\
-  \stdvr &\defeq \lambda \; W. 
-              \begin{aligned}[t]
-  & \{ \npair{i} \mid i \in \ints \} 
-    \union \\
-         & \{ \npair{\stdcap[\noperm] }  \} 
-           \union \\
-         & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
-         & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
-         & \{ \npair{\stdcap[\exec]} \mid \\
-         & \quad\readCond{n,W,\start,\addrend} \land \\
-         & \quad \execCond{n,W,\start,\addrend,\exec} \}
-           \union \\
-           % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
-         & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
-         & \{ \npair{\stdcap[\rwx]} \mid \\
-         & \quad \writeCond{n,W,\start,\addrend} \land \\
-         & \quad \execCond{n,W,\start,\addrend,\exec} \land \\
-         & \quad \execCond{n,W,\start,\addrend,\rwx} \}
-       \end{aligned}
-\end{align*}
-
-\begin{align*}
-  \observations \quad : & \quad  \Worlds \rightarrow \UPred{\Regs \times \HeapSegments} \\
-  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
-                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
-                           & \qquad \Rightarrow \exists W' \future W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
-\end{align*}
-Removed the condition $\npair[n-i]{\heap'(0)} \in \stdvr(W)$.\\
-
-The ``good observations'' are as follows
-\begin{itemize}
-\item The execution does not stop in $n$ steps. 
-\item The execution stops with a $\failed$.
-\item The execution halts in at most $n$ steps with a heap where the island that governs address 0 is satisfied by some heap segment.
-\end{itemize}
-Address 0 is the designated I/O address.
-\\ \lau{The two first conditions where implied by the last one, so they were removed.
-It turned out to cause troubles to special case on address 0, so now the entire heap has to satisfy a future world (which is a bit worrying as only the i/o address is observable).}
-
-Register-file relation:
-\begin{align*}
-  \stdrr \quad : & \quad \Worlds \monfun \UPred{\Regs} \\
-  \stdrr(W) \defeq & \{ \npair{\reg} \mid \\
-                    & \quad \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
-                    & \qquad  \npair{\reg(r)} \in \stdvr(W) \}
-\end{align*}
-
-``Continuation'' relation:
-\begin{align*}
-  \stdkr \quad : & \quad  \Worlds \monfun \UPred{\Words} \\%TODO: Check if this is still monotone
-  \stdkr(W) \defeq & \{ \npair{c} \mid \npair{c} \in \stdvr(W) \land \\
-                   & \quad \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
-                     \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(W') \ldotp \\
-                   & \qquad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
-\end{align*}
-We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.
-
-Normally, it would just be required for the return value to be in the value relation, so the continuation is invoked with this value.
-Here everything that is in the register file is available when the continuation is invoked, so we do not special case on $r_1$ (which is the register that in the calling convention we use should contain the return value).
-
-``Expression'' relation:
-\begin{align*}
-  \stder \quad : & \quad \Worlds \rightarrow \UPred{\Words} \\%TODO: Check whether this needs to be monotone
-  \stder(W) \defeq & \{ \npair{\pc} \mid \\
-                   & \quad \forall n' \leq n \ldotp\\
-                   & \qquad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
-                   & \qquad \quad \forall \npair[n']{c} \in \stdkr(W) \ldotp \\
-                   & \qquad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
-                   & \qquad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
-\end{align*}
-\lau{Added $n' \leq n$ to be able to prove Lemma~\ref{lem:stder-dc}}
-
+\subsection{Fundamental Theorem of Logical Relations}
 \begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline

### 2016-10-20 capability-machine.tex
Added lemmas from handwritten notes and fixed typos

@@
                                                               \text{if for $i \in \{1,2\}$}\\
-                                                              \quad\text{$n_i = \rv_i$ or $n_i = \memreg{\rv_i}$}\\
+                                                              \quad\text{$n_i = \rv_i$ or $n_i = \memreg(\rv_i)$}\\
                                                               \quad\text{and in either case $n_i \in \ints$}
@@
 \]
+$\xi$ is a morphism in the category $\CatP$ (the category of preordered \cofes{}), so it is non-expansive and monotone.
+
 We now define regions as
@@
 \end{align*}
-
 \end{definition}
+
+\begin{lemma}[Heap satisfaction uniformity]
+\label{lem:heapsat-dc}
+  \begin{align*}
+    & \forall n,n',\hs, W \ldotp\\
+    & \quad \heapSat[\hs]{n}{W} \land n' \leq n \Rightarrow \heapSat[\hs]{n'}{W}
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Heap satisfaction non-expansive]
+\label{lem:heapsat-ne}
+  \begin{align*}
+    & \forall n,\hs, W, W' \ldotp\\
+    & \quad \heapSat[\hs]{n}{W} \land W \nequal W' \Rightarrow \heapSat[\hs]{n}{W'}
+  \end{align*}
+\end{lemma}
 
@@
   \begin{align*}
-    &\forall \stdvr, n' < n \ldotp \forall W \ldotp \forall \reg \ldotp\\
+    &\forall \stdvr, n' \leq n \ldotp \forall W \ldotp \forall \reg \ldotp\\
     &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair[n']{\reg} \in \stdrr(\stdvr)(W)
@@
   \begin{align*}
-    &\forall \stdvr \ldotp \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp\\
+    &\forall \stdvr \ldotp \forall n' \leq n \ldotp \forall W \ldotp \forall c \ldotp\\
     &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair[n']{c} \in \stdkr(\stdvr)(W)
@@
   \begin{align*}
-    & \forall \stdvr \ldotp \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp\\
+    & \forall \stdvr \ldotp \forall n' \leq n \ldotp \forall W \ldotp \forall \pc \ldotp\\
     & \quad \npair{\pc} \in \stder(\stdvr)(W) \Rightarrow \npair[n']{\pc} \in \stder(\stdvr)(W) 
@@
   \iota_{\var{base},\var{end}} & \defeq \lambda \; \stdvr \ldotp ((\var{base},\var{end}),=,H_{\var{std}} (\stdvr)) \\ & \\
-  H_{\var{std}} & : (\Worlds \monnefun \UPred{\Words}) \nefun \States \nefun \Worlds \monnefun \UPred{\HeapSegments} \\
-  H_{\var{std}} \; \stdvr \; (\var{base},\var{end})\; W &  \defeq \left\{ \npair{\hs} \middle|
+  H_{\var{std}} & : (\Worlds \monnefun \UPred{\Words}) \nefun \States \nefun \Wor \monnefun \UPred{\HeapSegments} \\
+  H_{\var{std}} \; \stdvr \; (\var{base},\var{end})\; \hat{W} &  \defeq \left\{ \npair{\hs} \middle|
                                                \begin{aligned}
                                                  & \dom(\hs) = [\var{base},\var{end}] \land \\
-                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; W)
+                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; \hat{W})
                                                \end{aligned} \right\}
@@
 
+\begin{lemma}[$H_{\var{std}}$ is uniform]
+  \label{lem:std-reg-dc}
+  \begin{align*}
+    & \forall \stdvr, s=(\start,\addrend), \hat{W}, n', n, \hs \ldotp \\
+    & \quad n' \leq n \land \npair{\hs} \in H_{\var{std}} \; \stdvr \; s \; \hat{W} \\
+    & \qquad \Rightarrow \npair[n']{\hs} \in H_{\var{std}} \; \stdvr \; s \; \hat{W}
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[$H_{\var{std}}$ is non-expansive in state]
+  \label{lem:std-reg-ne-state}
+  \begin{align*}
+    & \forall \stdvr, (\start,\addrend), (\start',\addrend'), n \ldotp \\
+    & \quad  (\start,\addrend) \nequal (\start',\addrend') \\
+    & \qquad \Rightarrow H_{\var{std}} \; \stdvr \; (\start,\addrend) \nequal H_{\var{std}} \; \stdvr \; (\start',\addrend')
+  \end{align*}
+\end{lemma}
+
 \begin{lemma}[$\iota_{\start,\addrend}$ is non-expansive in the value relation]
@@
 \begin{align*}
-  \readCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
-  \readCond{}(\stdvr)(\start,\addrend,W) & =        
+  & \readCond{} : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
+  & \readCond{}(\stdvr)(\start,\addrend,W) =        
   \begin{aligned}[t]
@@
  & \\
-  \writeCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
-  \writeCond{}(\stdvr)(\start,\addrend, W) & = 
+  & \writeCond{} : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
+  & \writeCond{}(\stdvr)(\start,\addrend, W) = 
   \begin{aligned}[t]
@@
  & \\
-  \execCond{} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Perms \times \Worlds) \monnefun \DCPred{\nats} \\
-  \execCond{}(\stdvr)(\start,\addrend,\perm,W) & = 
+  & \execCond{} : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Perms \times \Worlds) \monnefun \DCPred{\nats} \\
+  & \execCond{}(\stdvr)(\start,\addrend,\perm,W) = 
   \begin{aligned}[t]
@@
  & \\
-  \entryCond{} & : \;  (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^3 \times \Worlds) \monnefun \DCPred{\nats} \\
-  \entryCond{}(\stdvr)(\start,\addrend,\addr,W) & =
+  & \entryCond{} : \;  (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^3 \times \Worlds) \monnefun \DCPred{\nats} \\
+  & \entryCond{}(\stdvr)(\start,\addrend,\addr,W) =
   \begin{aligned}[t]
@@
   \begin{align*}
-    &\forall \stdvr, n,W,W',\start, \addrend \ldotp \\
-    &\quad  (\start,\addrend,W') \future (\start,\addrend,W) \\
-    &\qquad \Rightarrow \readCond{}(\stdvr)(\start,\addrend,W') \supseteq \readCond{}(\stdvr)(\start,\addrend,W)
+    &\forall \stdvr, n,W,W',\start, \addrend, \start',\addrend' \ldotp \\
+    &\quad  (\start',\addrend',W') \future (\start,\addrend,W) \\
+    &\qquad \Rightarrow \readCond{}(\stdvr)(\start',\addrend',W') \supseteq \readCond{}(\stdvr)(\start,\addrend,W)
   \end{align*}
@@
   \begin{align*}
-    & \forall \stdvr, b,e, n, W, W' \ldotp \\
-    & \quad (b,e,W) \nequal (b,e,W') \Rightarrow \\
-    & \qquad \readCond{}(\stdvr)(b,e,W) \nequal \readCond{}(\stdvr)(b,e,W')\\
+    & \forall \stdvr, \start, \addrend, \start', \addrend', n, W, W' \ldotp \\
+    & \quad (\start, \addrend,W) \nequal (\start', \addrend',W') \Rightarrow \\
+    & \qquad \readCond{}(\stdvr)(\start, \addrend,W) \nequal \readCond{}(\stdvr)(\start', \addrend',W')\\
   \end{align*}
@@
   \begin{align*}
-    &\forall \stdvr, n,W,W',\start, \addrend \ldotp \\
-    &\quad  (\start,\addrend,W') \future (\start,\addrend,W) \\
-    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start,\addrend,W') \supseteq \writeCond{}(\stdvr)(\start,\addrend,W)
+    &\forall \stdvr,W,W',\start, \addrend, \start',\addrend' \ldotp \\
+    &\quad  (\start',\addrend',W') \future (\start,\addrend,W) \\
+    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start',\addrend',W') \supseteq \writeCond{}(\stdvr)(\start,\addrend,W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Read-write condition non-expansive in world]
+  \label{lem:wc-ne-world}
+  \begin{align*}
+    &\forall \stdvr, n,W,W',\start, \addrend, \start',\addrend' \ldotp \\
+    &\quad  (\start',\addrend',W') \nequal (\start,\addrend,W) \\
+    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start,\addrend,W) \nequal \writeCond{}(\stdvr)(\start',\addrend',W')
   \end{align*}
@@
   \begin{align*}
-    &\forall \stdvr,n,W, W',\start, \addrend, \perm \ldotp \\
-    &\quad  (\start,\addrend,\perm,W') \future (\start,\addrend,\perm,W) \\
-    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\perm,W') \supseteq \execCond{}(\stdvr)(\start,\addrend,\perm,W)
+    &\forall \stdvr,n,W, W',\start, \addrend, \perm,\start', \addrend', \perm' \ldotp \\
+    &\quad  (\start',\addrend',\perm',W') \future (\start,\addrend,\perm,W) \\
+    &\qquad \Rightarrow \execCond{}(\stdvr)(\start',\addrend',\perm',W') \supseteq \execCond{}(\stdvr)(\start,\addrend,\perm,W)
   \end{align*}
@@
   \begin{align*}
-    &\forall \stdvr, W_1, W_2,n,\start,\addrend,\perm \ldotp \\
-    &\quad (\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2) \Rightarrow \\
-    &\qquad \Rightarrow \execCond{}(\stdvr)(b,e,p,W_1) \nequal \execCond{}(\stdvr)(b,e,p,W_2)
+    &\forall \stdvr, W, W',n,\start,\addrend,\perm,\start',\addrend',\perm' \ldotp \\
+    &\quad (\start,\addrend,\perm,W) \nequal (\start',\addrend',\perm',W') \Rightarrow \\
+    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\perm,W) \nequal \execCond{}(\stdvr)(\start',\addrend',\perm',W')
   \end{align*}
@@
   \begin{align*}
-    &\forall \stdvr,n,W, W',\start, \addrend, \addr \ldotp \\
-    &\quad  (\start,\addrend,\addr,W') \future (\start,\addrend,\addr,W) \\
-    &\qquad \Rightarrow \entryCond{}(\stdvr)(\start,\addrend,\perm,W') \supseteq \entryCond{}(\stdvr)(\start,\addrend,\perm,W)
+    &\forall \stdvr,W, W',\start, \addrend, \addr, \start', \addrend', \addr' \ldotp \\
+    &\quad  (\start',\addrend',\addr',W') \future (\start,\addrend,\addr,W) \\
+    &\qquad \Rightarrow \entryCond{}(\stdvr)(\start',\addrend',\perm',W') \supseteq \entryCond{}(\stdvr)(\start,\addrend,\perm,W)
   \end{align*}
@@
   \begin{align*}
-    &\forall \stdvr, W_1, W_2,n,\start,\addrend,\addr \ldotp \\
-    &\quad (\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2) \Rightarrow \\
-    &\qquad \Rightarrow \execCond{}(\stdvr)(b,e,p,W_1) \nequal \execCond{}(\stdvr)(b,e,p,W_2)
+    &\forall \stdvr,n, W, W',\start,\addrend,\addr, \start', \addrend', \addr' \ldotp \\
+    &\quad (\start,\addrend,\perm,W) \nequal (\start',\addrend',\perm',W') \Rightarrow \\
+    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\addr,W) \nequal \execCond{}(\stdvr)(\start',\addrend',\addr',W')
   \end{align*}
@@
 \end{align*}
+
+\begin{lemma}[Value relation uniformity]
+  \begin{align*}
+    & \forall V, W, n, n' \ldotp \\
+    & \quad n' \leq n \land n \in \stdvr(V)(W) \\
+    & \qquad \Rightarrow n' \in \stdvr(V)(W)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Value relation non-expansive in $V$]
+  \begin{align*}
+    & \forall V, V', n \ldotp \\
+    & \quad V \nequal V' \Rightarrow \stdvr(V) \nequal \stdvr(V')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Value relation monotone in world]
+  \begin{align*}
+    & \forall V, W, W' \ldotp \\
+    & W' \future W \Rightarrow \stdvr(V)(W') \supseteq \stdvr(V)(W)
+  \end{align*} 
+\end{lemma}
+
+\begin{lemma}[Value relation non-expansive in world]
+  \begin{align*}
+    & \forall V, n, W, W' \ldotp \\
+    & W \nequal W' \Rightarrow \stdvr(V)(W) \nequal \stdvr(V)(W')
+  \end{align*} 
+\end{lemma}
+
+\begin{lemma}[Value relation contractive]
+  \begin{align*}
+    & \forall V, V', n \ldotp \\
+    & \quad V \nequal V' \Rightarrow \stdvr(V) \nequal[n+1] \stdvr(V')
+  \end{align*}
+\end{lemma}
+
 Using the Banach fixed-point theorem, we can find a fixed point for $\stdvr$ We will use this fixed point in the next section to define the logical relation we use in practice.
@@
 The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notive that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
-\end{proof}
-
-\begin{lemma}[Heap satisfaction uniformity]
-  \begin{align*}
-    & \forall n,n',\heap, W \ldotp\\
-    & \quad \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  %TODO
-\end{proof}
-
-\begin{lemma}[Heap satisfaction non-expansive]
-  \begin{align*}
-    & \forall n,\heap, W, W' \ldotp\\
-    & \quad \heapSat{n}{W} \land W \nequal W' \Rightarrow \heapSat{n}{W'}
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  %TODO
 \end{proof}

### 2016-10-20 capability-machine.tex
Added two missing lemmas (not yet proven) and some labels

@@
 
+\begin{lemma}[Register relation non-expansive in worlds]
+  \label{lem:reg-ne-worlds}
+  \begin{align*}
+    &\forall \stdvr, n \ldotp \forall W, W' \ldotp \\
+    &\quad W \nequal W' \\
+    &\qquad \Rightarrow \stdrr(\stdvr)(W) \nequal \stdrr(\stdvr)(W')
+  \end{align*}
+\end{lemma}
+
 \begin{lemma}[Register relation non-expansive in value relation]
@@
     &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair{c} \in \stdkr(\stdvr)(W')
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Continuation relation non-expansive in worlds]
+  \label{lem:stdkr-ne-worlds}
+  \begin{align*}
+    &\forall \stdvr, n \ldotp \forall W, W' \ldotp\\
+    &\quad W \nequal W' \\
+    &\qquad \Rightarrow \stdkr(\stdvr)(W) \nequal \stdkr(\stdvr)(W')
   \end{align*}
@@
 \end{lemma}
+
 \begin{lemma}[Expression relation non-expansive in world]
@@
 \begin{align*}
-  \forall \stdvr \ldotp \forall W_1 \nequal W_2 \ldotp \stder(\stdvr)(W_1) \nequal \stder(\stdvr)(W_2)
+  & \forall \stdvr \ldotp \forall W_1 \nequal W_2 \ldotp \stder(\stdvr)(W_1) \nequal \stder(\stdvr)(W_2)
 \end{align*}
@@
 \begin{lemma}[Value relation uniformity]
+\label{lem:stdvr-dc}
   \begin{align*}
@@
 \begin{lemma}[Value relation non-expansive in $V$]
+\label{lem:stdvr-ne-V}
   \begin{align*}
@@
 
-\begin{lemma}[Value relation monotone in world]
+\begin{lemma}[Value relation monotone in worlds]
+\label{lem:stdvr-mono-world}
   \begin{align*}
@@
 
-\begin{lemma}[Value relation non-expansive in world]
+\begin{lemma}[Value relation non-expansive in worlds]
+\label{lem:stdvr-ne-worlds}
   \begin{align*}
@@
 \begin{lemma}[Value relation contractive]
+\label{lem:stdvr-con}
   \begin{align*}
@@
 
-Using the Banach fixed-point theorem, we can find a fixed point for $\stdvr$ We will use this fixed point in the next section to define the logical relation we use in practice.
+Seeing as Lemma~\ref{lem:stdvr-con} gives us that $\stdvr$ is contractive, the Banach fixed-point theorem gives us a fixed-point for $\stdvr$.
+We will use this fixed point in the next section to define the logical relation we use in practice.
 

### 2016-10-21 capability-machine.tex
restructured proof section to reflect new presentation

@@
 \subsection{Worlds}
+\label{subsec:worlds}
 Assume preordered \cofe{}\ $\Wor$ and isomorphism $\xi$ such that:
@@
 To define future worlds and regions, We use the ordering inherited from the preordered \cofes{}.
+\begin{definition}[Future regions]
+For regions $(s_2,\phi_2,H_2), (s_1,\phi_1,H_1) \in \Regions$
+  \begin{align*}
+ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) &&\text{ iff } & &
+(\phi_1,H_1) = (\phi_2,H_2) \text{ and } (s_1,s_2) \in \phi_2
+  \end{align*}
+\end{definition}
+
 \begin{definition}[Future worlds]
@@
 
-\begin{definition}[Future regions]
-For regions $(s_2,\phi_2,H_2), (s_1,\phi_1,H_1) \in \Regions$
-  \begin{align*}
- (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) &&\text{ iff } & &
-(\phi_1,H_1) = (\phi_2,H_2) \text{ and } (s_1,s_2) \in \phi_2
-  \end{align*}
-\end{definition}
+\begin{lemma}[Future worlds transitive]
+\label{lem:future-worlds-trans}
+  \begin{align*}
+    & \forall W, W', W'' \ldotp \\
+    & \quad  W'' \future W' \land W' \future W \Rightarrow W'' \future W
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}
+\label{lem:nequal-and-future-world}
+  \begin{align*}
+    & \forall W_1,W_2,W_1' \ldotp \\
+    & \quad W_1 \nequal W_2 \land W_1' \future W_1 \Rightarrow \\
+    & \qquad \exists W_2' \ldotp W_2' \future W_2 \land W_1' \nequal W_2'
+  \end{align*}
+\end{lemma}
 
@@
 \subsubsection{Continuation Relation}
-\label{subsubsection:continuation-relation}
+\label{subsubsec:continuation-relation}
 The continuation relation is used in the definition of the expression relation.
@@
 \subsubsection{Expression Relation}
-\label{subsubsection:expression-relation}
+\label{subsubsec:expression-relation}
 The expression relation is defined as follows:
@@
 
+\begin{lemma}[Write condition implies read condition]
+\label{lem:wc-implies-rc}
+  \begin{align*}
+    &\forall n, W, \start, \addrend \ldotp\\ 
+    &\quad \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
+  \end{align*}
+\end{lemma}
+
 \subsubsection{Value Relation}
@@
 
-\subsection{Well-foundedness lemmas}
-\begin{lemma}[Future worlds transitive]
-  \begin{align*}
-    & \forall W, W', W'' \ldotp \\
-    & \quad  W'' \future W' \land W' \future W \Rightarrow W'' \future W
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  %TODO
-\end{proof}
-
-\begin{lemma}
-\label{lem:nequal-and-future-world}
-  \begin{align*}
-    & \forall W_1,W_2,W_1' \ldotp \\
-    & \quad W_1 \nequal W_2 \land W_1' \future W_1 \Rightarrow \\
-    & \qquad \exists W_2' \ldotp W_2' \future W_2 \land W_1' \nequal W_2'
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Assume $W_1 \nequal W_2$ and $W_1' \future W_2'$. Define $W_2'$ to have $\dom(W_2')=\dom(W_1')$ where
-\[
-  W_2' \; r = \left\{
-  \begin{aligned}
-    & (s_1',\phi_2,H_2) & &\text{for } j \in \dom(W_2) \text{ and } W_2 \; j = (s_2,\phi_2,H_2) \\
-    &                  & &\quad \text{ and } W_1' \; j = (s_1',\phi',H_1') \\
-    & W_1' \; r        & &\text{otherwise}
-  \end{aligned} 
-  \right.
-\]
-
-First we show $W_1' \nequal W_2'$.
-The domains are defined to be equal, so that is satsfied.
-Let $r \in \dom(W_1')$ and show $W_1' \; r \nequal W_2' \; r$.
-First consider the case where $r \in \dom(W_2)$.
-Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$.
-We need to show
-\begin{align*}
-  s_1' & \nequal s_1' \\
-  \phi_1' & \nequal \phi_2 \\
-  H_1' & \nequal H_2
-\end{align*}
-The first one is trivial.
-For the second one we use that we know $\phi_1 \nequal \phi_2$ and $\phi_1 = \phi_1'$ respectively from the $n$-equality assumption and future world assumption.
-Similarly, the second one follows from $H_1 \nequal H_2$ and $H_1 = H_1'$ - again respectively from the $n$-equality assumption and the future world assumption.
-For the case where $r \not\in \dom(W_2)$, we need to show $W_1' \; r \nequal W_1' \; r$ which follows trivially from reflexivity.
-
-Next we show $W_2' \future W_2$.
-We know that $\dom(W_2') = \dom(W_1')$, $\dom(W_1') \supseteq \dom(W_1)$ (from future world assumption), and $\dom(W_1) = \dom(W_2)$ (from $n$-equality assumption).
-All this can easily be put together to get $\dom(W_2') \supseteq \dom(W_2)$.
-Now let $r\in \dom(W_2)$ and assume Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$, then we need to show the following:
-\begin{align*}
-  &(s_2,s_1') \in \phi_2 \\
-  &\phi_2 = \phi_2 \\
-  &H_2 = H_2
-\end{align*}
-Here the two last are trivial.
-The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notive that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-dc}]
-  Follows directly from definition: Let $\stdvr$, $n$, $n'$, $\start$, and  $\addrend$ be given.
-Assume $n \in \readCond{}(\stdvr)(\start,\addrend,W)$ and $n' \leq n$.
-If $n=n'$, then the result follows trivially, as it is exactly our assumption.
-If $n' < n$, then by assumption we have a region name $r$, interval $[\start',\addrend'] \supseteq [\start,\addrend]$, and the assumption $W(r) \nsubeq[n-1] \iota_{\start',\addrend'}$  given by the read condition assumption.
-Picking $r$ and $[\start',\addrend']$, we need to show  $W(r) \nsubeq[n'-1] \iota_{\start',\addrend'}$ which follows from downwards closure of $\nsubeq[n-1]$.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-mono-world}]
-Let $\stdvr$, $n$,$W$, $W'$, $b$, and $e$ be given where $W' \future W$.
-Let $n \in \readCond{}(\stdvr)(\start,\addrend,W)$, this gives an $r$ and $[b',e'] \supseteq [b,e]$ such that $W(r) \nsubeq \iota_{b',e'} (r) = ((b',e'),=,H_{\var{std}})$.
-This gives
-\begin{align*}
-  &W(r).\phi \equiv = \\
-  &W(r).s    = (b',e') \\
-  &\forall \hat{W} \ldotp W(r).H\; (b',e') \; \hat{W} \nsubeq[n-1] H_{\var{std}}\; (b',e') \; \hat{W}
-\end{align*}
-$W' \future W$ gives
-\begin{align*}
-  &W'(r).\phi = W(r).\phi \\
-  &(W'(r).s,W(r).s) \in W(r).\phi \\
-  &W'(r).H = W(r).H \\
- \end{align*}
-Combining these two sets of properties line for line, we get the following for $W'$:
-\begin{align*}
-  & W'(r).\phi \equiv = \\
-  & W'(r) = (b',e') \\
-  &\forall \hat{W} \ldotp W'(r).H\; (b',e') \; \hat{W} \nsubeq[n-1] H_{\var{std}}\; (b',e') \; \hat{W}
-\end{align*}
-These are exactly the properties needed to get $W'(r) \nsubeq \iota_{b',e'} (r)$ which is exactly the result we want.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-ne-world}]
-Assume 
-\[
-  (b,e,W) \nequal (b,e,W')
-\]
-which implies $W \nequal W'$. Now let $(n,\hs) \in \readCond{}(\stdvr)(b,e,W)$ be given. This gives us $r$ and $[b',e'] \supseteq [b,e]$ such that 
-\[
-  W(r) \nsubeq[n-1] \iota_{b',e'}(\stdvr)
-\]
-Using $r$ and $[b',e']$, we need to show
-\[
-  W'(r) \nsubeq[n-1] \iota_{b',e'}(\stdvr)
-\]
-Let $W(r) = (s,\phi,H)$, $W(r) = (s',\phi',H')$, and $\iota_{b',e'} = (s_\iota,\phi_\iota,H_\iota)$.
-From the $n$-equality of our worlds, we get $W(r) \nequal W'(r)$. From this we conclude
-\[
-  (s',\phi') = (s,\phi) = (s_\iota,\phi_\iota)
-\]
-It remains to show
-\[
-  \forall \hat{W} \in \Wor \ldotp H' \; s \; W \nequal[n-1] H_\iota \; s \; W
-\]
-Given $\hat{W}$ we know from $n$-equality of the worlds
-\[
-   H' \; s \; \hat{W} \nequal H \; s \; \hat{W}
-\]
-Given $(k,\hs) \in H' \; s \; W$ where $k < n$ and $\hs$ is a heap segment, by above we have $(k,\hs) \in H \; s \; W$ and from the $\nsubeq[n-1]$ assumption, we get $(k,\hs) \in H_\iota \; s_\iota \; W$.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:wc-dc}]
-  Follows directly from the definition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:wc-mono-world}]
-  Similar to the proof of the monotonicity of the read condition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-dc}]
-  Follows directly from definition: Let $\stdvr$, $n$, $n'$, $W$, $\start$, and $\addrend$ be given.
-Assume $n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W)$ and $n' \leq n$.
-If $n' = n$, then the result follows directly from assumption.
-If $n' < n$, then let $n'' < n'$, $W' \future W$, and $a \in [\start,\addrend]$ be given.
-Using that we by transitivity have $n'' < n$, we get the desired result from assumption.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-mono-world}]
-  Follows directly from definition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-ne-world}]
-  Assume $(\start,\addrend,\perm,W_1) \nequal (\start,\addrend,\perm,W_2)$ and thus $W_1 \nequal W_2$. Given $n'<n$, $W_1' \future W_1$, and $\addr$, show
-\[
-  \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W_1')
-\]
-by lemma~\ref{lem:nequal-and-future-world} there must exist $W_2'$ such that $W_2' \future W_2$ and $W_2' \nequal W_1'$. By assumption $n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W_2)$ we have
-\[
-\npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W_2')
-\]
-and by non-expansiveness of $\stder(\stdvr)$ (lemma~\ref{lem:stder-ne-worlds}), we get the desired result.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-dc}]
-  Like the $\execCond{}$ uniformity proof.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-mono-world}]
-  Follows directly from definition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-ne-world}]
-  This proof goes like the proof of $\execCond{}$ being non-expansive in the second argument (lemma~\ref{lem:ec-ne-world}).
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:wc-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-con}]
-  %TODO
-\end{proof}
-
-\begin{lemma}[Write condition implies read condition]
-\label{lem:wc-implies-rc}
-  \begin{align*}
-    &\forall n, W, \start, \addrend \ldotp\\ 
-    &\quad \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
-  \end{align*}
-\end{lemma}[Proof of lemma~\ref{lem:wc-implies-rc}]
-\begin{proof}
-  Follows directly from the definition.
-\end{proof}
-
-\begin{lemma}[Value relation uniformity]
-\label{lem:stdvr-uniform-step}
-  \begin{align*}
-    &\forall n' < n \ldotp \forall W \ldotp \forall w \ldotp\\
-    &\quad \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Follows from the uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$.
-\end{proof}
-
-\begin{lemma}[Value relation monotone in worlds]
-  \begin{align*}
-    &\forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp\\ 
-    &\quad \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  Follows from uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$ in the worlds.
-That is Lemma~\ref{lem:rc-mono-world},~\ref{lem:wc-mono-world},~\ref{lem:ec-mono-world}, and~\ref{lem:en-mono-world}.
-\end{proof}
-
-\begin{lemma}[Value relation contractive]
-  \begin{align*}
-    & \forall n \ldotp \forall W \in \Worlds \forall w \ldotp \\
-    & \quad \npair[n]{w} \in \stdvr(W) \Rightarrow \\
-    & \qquad \npair[n+1]{w} \in \stdvr(W)
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:obs-dc}]
-  Let $n' < n$, $W$, $\reg$, and $\hs$ be given.
-Assume $\npair{(\reg,\hs)} \in \observations(W)$.
-Let $\heap_f$, $heap'$, $i \leq n'$ be given and assume $(\reg, \hs \uplus \heap_f) \step[i] (\halted, \heap')$.
-By assumption, we have a $W' \future W$ and $\hs'$ such that
-  \begin{align}
-&\heap' = \hs' \uplus \heap_f \label{lem:eq:stdvr1}\\
-&\heapSat[\hs']{n-i}{W'}\label{lem:eq:stdvr2}
-  \end{align}. 
-Using $W'$ and $\hs'$ as existential witnesses, we already have Equation~\ref{lem:eq:stdvr1} as the first necessary condition and from the above heap satisfaction along with heap satisfaction being uniform in $n$, we get $\heapSat[\hs']{n'-i}{W'}$.
-These are the two conditions necessary to get $\npair[n']{(\reg,\hs)} \in \observations(W)$.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:obs-ne-worlds}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:reg-dc}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:reg-mono-worlds}]
-  This follows by monotonicity of $\stdvr$ in the worlds.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:reg-ne-vr}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdkr-dc}]
-Follows by using the assumption and transitivity of inequality.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdkr-mono-worlds}]
-We need to show two things: The continuation is in the value predicate in the future world and that continuing execution using the continuation is in the observation predicate.
-
-The first follows from the value predicate being monotone in the world.
-
-The second property follows by assumption and transitivity of future worlds.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdkr-ne-vr}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stder-dc}]
-  Follows from the assumption.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stder-ne-worlds}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stder-ne-vr}]
-    %TODO
-\end{proof}
 \section{Other examples and applications}
@@
 \begin{appendices}
-\section{Definitions and Proofs related to Section~\ref{subsec:recursive-dom-eq}}
-\end{appendices}
+
+\section{Proofs of Lemmas in Section~\ref{sec:logical-relation}}
+\subsection{Definitions and Proofs related to Section~\ref{subsec:recursive-dom-eq}}
 
@@
 \end{proof}
-
-\section{Proof of Lemmas in Section~\ref{subsec:further-standard-regions}}
+\subsection{Proofs of Lemmas in Section~\ref{subsec:worlds}}
+%TODO All proofs written by hand, typeset them.
+\begin{proof}[Proof of lemma~\ref{lem:nequal-and-future-world}]
+  Assume $W_1 \nequal W_2$ and $W_1' \future W_2'$. Define $W_2'$ to have $\dom(W_2')=\dom(W_1')$ where
+\[
+  W_2' \; r = \left\{
+  \begin{aligned}
+    & (s_1',\phi_2,H_2) & &\text{for } j \in \dom(W_2) \text{ and } W_2 \; j = (s_2,\phi_2,H_2) \\
+    &                  & &\quad \text{ and } W_1' \; j = (s_1',\phi',H_1') \\
+    & W_1' \; r        & &\text{otherwise}
+  \end{aligned} 
+  \right.
+\]
+
+First we show $W_1' \nequal W_2'$.
+The domains are defined to be equal, so that is satsfied.
+Let $r \in \dom(W_1')$ and show $W_1' \; r \nequal W_2' \; r$.
+First consider the case where $r \in \dom(W_2)$.
+Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$.
+We need to show
+\begin{align*}
+  s_1' & \nequal s_1' \\
+  \phi_1' & \nequal \phi_2 \\
+  H_1' & \nequal H_2
+\end{align*}
+The first one is trivial.
+For the second one we use that we know $\phi_1 \nequal \phi_2$ and $\phi_1 = \phi_1'$ respectively from the $n$-equality assumption and future world assumption.
+Similarly, the second one follows from $H_1 \nequal H_2$ and $H_1 = H_1'$ - again respectively from the $n$-equality assumption and the future world assumption.
+For the case where $r \not\in \dom(W_2)$, we need to show $W_1' \; r \nequal W_1' \; r$ which follows trivially from reflexivity.
+
+Next we show $W_2' \future W_2$.
+We know that $\dom(W_2') = \dom(W_1')$, $\dom(W_1') \supseteq \dom(W_1)$ (from future world assumption), and $\dom(W_1) = \dom(W_2)$ (from $n$-equality assumption).
+All this can easily be put together to get $\dom(W_2') \supseteq \dom(W_2)$.
+Now let $r\in \dom(W_2)$ and assume Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$, then we need to show the following:
+\begin{align*}
+  &(s_2,s_1') \in \phi_2 \\
+  &\phi_2 = \phi_2 \\
+  &H_2 = H_2
+\end{align*}
+Here the two last are trivial.
+The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notive that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:rc-dc}]
+  Follows directly from definition: Let $\stdvr$, $n$, $n'$, $\start$, and  $\addrend$ be given.
+Assume $n \in \readCond{}(\stdvr)(\start,\addrend,W)$ and $n' \leq n$.
+If $n=n'$, then the result follows trivially, as it is exactly our assumption.
+If $n' < n$, then by assumption we have a region name $r$, interval $[\start',\addrend'] \supseteq [\start,\addrend]$, and the assumption $W(r) \nsubeq[n-1] \iota_{\start',\addrend'}$  given by the read condition assumption.
+Picking $r$ and $[\start',\addrend']$, we need to show  $W(r) \nsubeq[n'-1] \iota_{\start',\addrend'}$ which follows from downwards closure of $\nsubeq[n-1]$.
+\end{proof}
+
+\subsection{Proofs of Lemmas in Section~\ref{subsec:logical-relation-setup}}
+%All the lemmas are in subsections.
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:observation-relation}}
+%TODO All proofs written by hand (but one), typeset them.
+\begin{proof}[Proof of lemma~\ref{lem:obs-dc}]
+  Let $n' < n$, $W$, $\reg$, and $\hs$ be given.
+Assume $\npair{(\reg,\hs)} \in \observations(W)$.
+Let $\heap_f$, $heap'$, $i \leq n'$ be given and assume $(\reg, \hs \uplus \heap_f) \step[i] (\halted, \heap')$.
+By assumption, we have a $W' \future W$ and $\hs'$ such that
+  \begin{align}
+&\heap' = \hs' \uplus \heap_f \label{lem:eq:stdvr1}\\
+&\heapSat[\hs']{n-i}{W'}\label{lem:eq:stdvr2}
+  \end{align}. 
+Using $W'$ and $\hs'$ as existential witnesses, we already have Equation~\ref{lem:eq:stdvr1} as the first necessary condition and from the above heap satisfaction along with heap satisfaction being uniform in $n$, we get $\heapSat[\hs']{n'-i}{W'}$.
+These are the two conditions necessary to get $\npair[n']{(\reg,\hs)} \in \observations(W)$.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:obs-ne-worlds}]
+  %TODO
+\end{proof}
+
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:register-file-relation}}
+%TODO All proof but one proof written by hand, typeset them.
+%TODO missing proof register relation ne in worlds.
+\begin{proof}[Proof of lemma~\ref{lem:reg-dc}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:reg-mono-worlds}]
+ %TODO,  This follows by monotonicity of $\stdvr$ in the worlds.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:reg-ne-vr}]
+  %TODO
+\end{proof}
+
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:continuation-relation}}
+%TODO All proof but one proof written by hand, typeset them.
+%TODO missing proof continuation relation ne in worlds.
+\begin{proof}[Proof of lemma~\ref{lem:stdkr-dc}]
+  % TODO, Follows by using the assumption and transitivity of inequality.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdkr-mono-worlds}]
+  % TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdkr-ne-vr}]
+  %TODO
+\end{proof}
+
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:expression-relation}}
+%TODO All proofs written by hand, typeset them.
+\begin{proof}[Proof of lemma~\ref{lem:stder-dc}]
+  %TODO,  Follows from the assumption.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stder-ne-worlds}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stder-ne-vr}]
+  %TODO
+\end{proof}
+
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:standard-region}}
+%TODO All proofs written by hand, typeset them.
 \begin{proof}[Proof of Lemma~\ref{lem:std-reg-mono-world}]
@@
 
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:capability-conditions}}
+%TODO All proofs written by hand, typeset them.
+
+\begin{proof}[Proof of lemma~\ref{lem:rc-mono-world}]
+%TODO
+\end{proof}
+\begin{proof}[Proof of lemma~\ref{lem:rc-ne-world}]
+%TODO
+\end{proof}
+\begin{proof}[Proof of lemma~\ref{lem:wc-dc}]
+%TODO, Follows directly from the definition.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:wc-mono-world}]
+%TODO, Similar to the proof of the monotonicity of the read condition.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:ec-dc}]
+%TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:ec-mono-world}]
+%TODO, Follows directly from definition.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:ec-ne-world}]
+%TODO 
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:en-dc}]
+  Like the $\execCond{}$ uniformity proof.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:en-mono-world}]
+  Follows directly from definition.
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:en-ne-world}]
+  This proof goes like the proof of $\execCond{}$ being non-expansive in the second argument (lemma~\ref{lem:ec-ne-world}).
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:rc-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:wc-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:ec-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:en-con}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:wc-implies-rc}]
+  Follows directly from the definition.
+\end{proof}
+
+\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:value-relation}}
+%TODO All proofs written by hand, typeset them.
+\begin{proof}[Proof of lemma~\ref{lem:stdvr-dc}]
+  Follows from the uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$.
+  %TODO show one case to illustrate the above, refer to the lemmas in question
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdvr-ne-V}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdvr-mono-world}]
+  Follows from monotonicity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$ in the worlds.
+That is Lemma~\ref{lem:rc-mono-world},~\ref{lem:wc-mono-world},~\ref{lem:ec-mono-world}, and~\ref{lem:en-mono-world}.
+  %TODO show one case to illustrate the above
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdvr-ne-worlds}]
+  %TODO
+\end{proof}
+
+\begin{proof}[Proof of lemma~\ref{lem:stdvr-con}]
+  %TODO
+\end{proof}
+
+\subsection{Proofs of Lemmas in Section~\ref{subsec:further-standard-regions}}
+%TODO all the proofs need to be done.
+
 \begin{proof}[Proof of Lemma~\ref{lem:std-z-region-wd}]
@@
 \end{proof}
+\end{appendices}
 

### 2016-10-24 capability-machine.tex
Added decode function for permissions
corrected mistak in operational semantics where encodePerm was used instead of decodePerm
Added new section for local capabilities and proposed operational semantics.

@@
 \newcommand{\addrend}{\var{end}}
+\newcommand{\pwlv}{\var{pwl}}
 \newcommand{\mem}{\var{mem}}
@@
 \newcommand{\encodePerm}{\plainfun{encodePerm}{}}
+\newcommand{\decodePerm}{\plainfun{decodePerm}{}}
 \newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}
@@
 \newcommand{\Perms}{\plaindom{Perm}}
+\newcommand{\Globals}{\plaindom{Global}}
+\newcommand{\PWLs}{\plaindom{PermitWriteLocal}}
 
@@
 \newcommand{\rwx}{\plainperm{rwx}}
+
+%Global/local
+\newcommand{\local}{\plainperm{local}}
+\newcommand{\glob}{\plainperm{global}}
+
+%PWLs
+\newcommand{\pwl}{\plaindom{pwl}}
+\newcommand{\nwl}{\plaindom{nwl}}
 
@@
 \begin{align*}
-\decode &:\Words \rightarrow \Instrs
+\decode &:\Words \fun \Instrs
 \end{align*}
@@
 \begin{align*}
-\encodePerm &: \Perms \rightarrow \ints
-\end{align*}
+\encodePerm &: \Perms \fun \ints
+\end{align*}
+Further, assume an inverse function, $\decodePerm$, that decodes permissions
+\[
+  \decodePerm : \ints \fun \Perms
+\]
 \begin{align*}
@@
                                                    \quad\text{and in either case $n \in \ints$}\\
-                                                   \quad\text{and $\var{newPerm} = \encodePerm(n)$}\\
+                                                   \quad\text{and $\var{newPerm} = \decodePerm(n)$}\\
                                                    \quad\text{and $\var{newPerm} \sqsubseteq \perm$}\\
@@
 
+\section{Adding Local capabilities}
+In this section, we add local capabilities and define a new logical relation for this setting.
+\lau{Should we do this in a new document?}
+The definitions are the same as before apart from the changes below:
+\begin{align*}
+  \PWLs & = 2 \\
+  \Globals & = 2 \\
+  \Caps & = \Perms \times \Globals \times \PWLs \times \Addrs \times \Addrs \times \Addrs
+\end{align*}
+That is the only new domains are the two element domains $\Globals$ which represents whether a capability is global or local and $\PWLs$ which represents whether a capability is permit write local or not.
+The $\Caps$ domain change to include both of these.
+The ordering of $\Globals$ is specified in Figure~\ref{fig:glob-hier} and the ordering of $\PWLs$ is specified in Figure~\ref{fig:pwl-hier}.
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+  \node[main node] (1) {$\glob$};
+  \node[main node] (2) [below of=1] {$\local$};
+
+  \path[every node/.style={font=\sffamily\small}]
+    (1) edge (2);
+\end{tikzpicture}
+\caption{Permission hierarchy}
+\label{fig:glob-hier}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+  \node[main node] (1) {$\pwl$};
+  \node[main node] (2) [below of=1] {$\nwl$};
+
+  \path[every node/.style={font=\sffamily\small}]
+    (1) edge (2);
+\end{tikzpicture}
+\caption{Permission hierarchy}
+\label{fig:pwl-hier}
+\end{figure}
+
+In order to define the operationtal semantics, we assume a decode function, $\decodePerm$ from
+
+\[
+  \decodePerm : \ints \fun \perm \times 2 \times 2
+\]
+
+There are two parts of the operational semantics that changes.
+The first one is $\store$ as it is the only instruction that allows one to write to the heap.
+It will now have an extra condition that makes sure that local capabilities are only written through capabilities that permit write local.
+The semantics of $\restrict$ also changes to allow $\glob$ capabilities to become $\local$ and $\pwl$ (permit write local) to become $\nwl$ (no write local).
+
+\renewcommand{\stdcap}[1][\perm]{\left(#1,g,\pwlv,\start,\addrend,\addr \right)}
+
+\begin{align*}
+   \sem{\store{\refheap{r_1}}{\refreg{r_2}}} & = 
+                                             \begin{cases}
+                                               \stdUpdatePc{\updateHeap{\addr}{\var{w}}} &
+                                               \arraycolsep=0pt
+                                               \begin{array}{l}
+                                                 \text{if }\memreg(r_1) = (\perm,\_,\pwlv,\_,\_,\addr) = \var{c} \\
+                                                 \quad\text{and }\writeAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
+                                                 \quad\text{and } \var{w} = \memreg(r_2) \\
+                                                 \quad\text{and if } \var{w} = (\_, \var{g}, \_, \_, \_, \_)\text{, then } \pwlv = \pwl
+                                               \end{array}\\
+                                               \failed & \text{otherwise }
+                                             \end{cases}\\
+  \sem{\restrict{\refreg{r_1}}{\rv_1}{\rv_2}}           & =
+                                             \begin{cases}
+                                               \stdUpdatePc{\updateReg{r_1}{\var{c}}}  &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}{l}
+                                                   \text{if $\memreg(\rv_1) = c'$}\\
+                                                   \quad\text{and $c' = \stdcap$}\\
+                                                   \quad\text{and either $\rv_2 = n$ or $\memreg(\rv_2) = n$}\\
+                                                   \quad\text{and in either case $n \in \ints$}\\
+                                                   \quad\text{and $(\perm', \var{g}',\pwlv') = \decodePerm(n)$}\\
+                                                   \quad\text{and $\perm' \sqsubseteq \perm$}\\
+                                                   \quad\text{and $\var{g}' \sqsubseteq \var{g}$}\\
+                                                   \quad\text{and $\pwlv'\sqsubseteq \pwlv$}\\
+                                                   \quad\text{and $c = (\perm',
+\var{g}', \pwlv',\start,\addrend,\addr)$}
+                                                 \end{array}\\
+                                               \failed                   & \text{otherwise}
+                                             \end{cases} 
+\end{align*}
+
 \section{Other examples and applications}

### 2016-10-25 capability-machine.tex
Started adding the local capability changes in the sections that were already written

--- 2016-10-21 capability-machine.tex
@@
 \newcommand{\lv}{\var{lv}}
+\newcommand{\gl}{g}
 \newcommand{\pc}{\mathit{pc}}
@@
 \newcommand{\perm}{\var{perm}}
+\newcommand{\permp}{\var{permPair}}
 \newcommand{\roll}{\var{roll}}
 \newcommand{\instr}{\var{instr}}
-
-\newcommand{\stdcap}[1][\perm]{\left(#1,\start,\addrend,\addr \right)}
+\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
 
@@
 \newcommand{\Perms}{\plaindom{Perm}}
+\newcommand{\Globals}{\plaindom{Global}}
 
@@
 \newcommand{\DCPred}[1]{\plaindom{P}^\downarrow(#1)}
+
+\newcommand{\Views}{\plaindom{View}}
 
@@
 \newcommand{\rwx}{\plainperm{rwx}}
+%PWL permissions
+\newcommand{\readwritel}{\plainperm{rwl}}
+\newcommand{\rwlx}{\plainperm{rwlx}}
+
+%Global/local
+\newcommand{\local}{\plainperm{local}}
+\newcommand{\glob}{\plainperm{global}}
 
@@
 \newcommand{\step}[1][]{\rightarrow_{#1}}
+
+%View
+\newcommand{\perma}{\mathrm{permanent}}
+\newcommand{\temp}{\mathrm{temporary}}
 
@@
 \end{flushright}
-$\RegName$ contains $\pcreg$, but is otherwise some undefined, finite set.
+\section{Domains and Notation}
 \begin{align*}
@@
 \Heaps &::= \Addrs \rightarrow \Words \\
-\Perms &::= \{ \noperm, \readonly, \readwrite, \exec, \entry, \rwx\} & & &
+\Perms &::= \{ \noperm, \readonly, \readwrite, \readwritel, \exec, \entry, \rwx, \rwlx\} & & &
 \Mems  &::= \Regs \times \Heaps \\
-\Caps  &::= \Perms \times \Addrs \times \Addrs \times \Addrs & & &
+\Caps  &::= (\Perms \times \Globals) \times \Addrs \times \Addrs \times \Addrs & & &
 \Confs &::= \Mems + \{\failed, \halted \times \Mems\} \\
-\HeapSegments &::= \Addrs \parfun \Words & & & &
-\end{align*}
-
+\HeapSegments &::= \Addrs \parfun \Words & & &
+\Globals & ::= \{\glob, \local \}
+\end{align*}
+Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capabaility is local or global.
+Two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities.
+They are otherwise the same as their non-''permit write local'' counterparts.
+
+The $\glob$ and $\local$ are ordered as shown in Figure~\ref{fig:glob-hier}.
 \begin{figure}[!h]
@@
   \begin{tikzpicture}[main node/.style={}]
-  \node[main node] (1) {$\rwx$};
+  \node[main node] (1) {$\glob$};
+  \node[main node] (2) [below of=1] {$\local$};
+
+  \path[every node/.style={font=\sffamily\small}]
+    (1) edge (2);
+\end{tikzpicture}
+\caption{Permission hierarchy}
+\label{fig:glob-hier}
+\end{figure}
+
+Things to note:
+\begin{itemize}
+\item $\RegName$ contains $\pcreg$, but is otherwise a sufficiently
+  large finite set.
+\item Table~\ref{tab:permission-list} describes what all the permissions grant acces to.
+\item Figure~\ref{fig:perm-hier} shows the ordering of the permissions, i.e, the elements of $\Perms$.
+\item Figure~\ref{fig:glob-hier} shows the ordering of $\local$ and $\glob$, i.e., the elements of $\Globals$.
+\item The ordering of $\Perms \times \Globals$ is pointwise.
+\end{itemize}
+
+\begin{table}[!h]
+  \centering
+  \begin{tabular}[!h]{r |  p{7cm} }
+  $\noperm$ & No permissions. Grants no permissions\\
+\hline
+  $\readonly$ & Read only. Grants read permission \\
+\hline
+  $\readwrite$ & Read-write. Grants read and write permissions. Storage of local capabilities prohibited. \\
+\hline
+  $\readwritel$ & read-write, permit write local. Grants read and write permissions. Storage of local capabilities granted. \\
+\hline
+  $\exec$ & Execute permission. Grants execute and read permissions.\\
+\hline
+  $\entry$ & Entry permission. This permission grants no access, but when jumped to, it Will turn into an $\exec$ permission.\\
+\hline
+  $\rwx$ & Read, write, execute permission. Grants read, write, and execute permissions. Storage of local capabilities prohibited. \\
+\hline
+  $\rwlx$ & Read, write, execute permission with permit write local. Grants read, write, and execute permissions. Storage of local capabilities possible.
+\end{tabular}
+
+\caption{The permissions in this capability system}
+\label{tab:permission-list}
+\end{table}
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (7) {$\rwlx$};
+    \node[main node] (8) [below left of=7] {$\readwritel$};
+    \node[main node] (1) [below right of=7] {$\rwx$};
     \node[main node] (2) [below right of=1] {$\exec$};
@@
   \path[every node/.style={font=\sffamily\small}]
+    (7) edge (8)
+    (7) edge (1)
+    (8) edge (4)
     (1) edge (2)
@@
 \end{tikzpicture}
+
 \caption{Permission hierarchy}
@@
 \begin{align*}
-\encodePerm &: \Perms \rightarrow \ints
-\end{align*}
+\encodePerm &: (\Perms \times \Globals) \fun \ints
+\end{align*}
+Further, assume an inverse function, $\decodePerm$, that decodes permissions
+\[
+  \decodePerm : \ints \fun (\Perms \times \Globals)
+\]
 \begin{align*}
@@
                                                                 \quad\text{and $\start \leq \addr < \addrend$}\\
-                                                                \quad\text{and $\perm \in \{ \exec,\rwx \}$ }
+                                                                \quad\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$ }
                                                               \end{array}\\
@@
                            \begin{cases}
-                             \true & \text{if } \perm \in \{ \rwx, \exec, \entry \} \\
+                             \true & \text{if } \perm \in \{ \rwx, \rwlx, \exec, \entry \} \\
                              \false & \text{otherwise}
@@
                            \begin{cases}
-                             \true & \text{if } \perm \in \{ \rwx, \exec, \readwrite, \readonly \} \\
+                             \true & \text{if } \perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \} \\
                              \false & \text{otherwise}
@@
                            \begin{cases}
-                             \true & \text{if } \perm \in \{ \rwx, \readwrite\} \\
+                             \true &
+                               \text{if } \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\} \\
                              \false & \text{otherwise}
@@
                                      \begin{cases}
-                                       (\exec,\start,\addrend,\addr) & \text{if $w = (\entry,\start,\addrend,\addr)$}\\
+                                       ((\exec,\gl),\start,\addrend,\addr) & \text{if $w = ((\entry,\gl),\start,\addrend,\addr)$}\\
                                        w & \text{otherwise} 
@@
                              \text{if $\memreg(\pcreg) = \stdcap$}\\
-                             \quad\text{and $\var{newPc} = (\perm,\start,\addrend,\addr + 1)$}\\
+                             \quad\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}\\
                            \end{array} \\
@@
                                                \begin{array}{l}
-                                                 \text{if }\memreg(r_2) = (\perm,\start,\addrend,\addr) = \var{c} \\
+                                                 \text{if }\memreg(r_2) = \stdcap = \var{c} \\
                                                  \quad\text{and }\readAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
@@
                                                \begin{array}{l}
-                                                 \text{if }\memreg(r_1) = (\perm,\start,\addrend,\addr) = \var{c} \\
+                                                 \text{if }\memreg(r_1) = \stdcap = \var{c} \\
                                                  \quad\text{and }\writeAllowed{\perm} \text{ and } \withinBounds{\var{c}} \\
-                                                 \quad\text{and }\var{w} = \memreg(r_2)
+                                                 \quad\text{and }\var{w} = \memreg(r_2)\\
+                                                 \quad\text{and if } \var{w} = ((\_,\local),\_,\_,\_) \text{,} \\
+                                                 \quad\text{ then } \perm \in \{\rwlx,\readwritel \}
                                                \end{array}\\
@@
                                                    \quad\text{and $\memreg(r_1) = \stdcap$}\\
-                                                   \quad\text{and $\var{c} = (\perm,\start,\addrend,\addr + n)$}
+                                                   \quad\text{and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$}
                                                  \end{array}\\
@@
                                                    \text{if $\memreg(\rv_1) = c'$}\\
-                                                   \quad\text{and $c' = \stdcap$}\\
+                                                   \quad\text{and $c' = \stdcap[\permp]$}\\
                                                    \quad\text{and either $\rv_2 = n$ or $\memreg(\rv_2) = n$}\\
                                                    \quad\text{and in either case $n \in \ints$}\\
-                                                   \quad\text{and $\var{newPerm} = \encodePerm(n)$}\\
-                                                   \quad\text{and $\var{newPerm} \sqsubseteq \perm$}\\
+                                                   \quad\text{and $\var{newPermPair} = \decodePerm(n)$}\\
+                                                   \quad\text{and $\var{newPermPair} \sqsubseteq \permp$}\\
                                                    \quad\text{and $c = (\var{newPerm},\start,\addrend,\addr)$}
@@
 \end{align*}
+\lau{Dominique, apparently there was a reason I had not added the new condition in store to $\writeAllowed{}$.
+The new condition depends on a case distinction on $w$.
+It is only if $w$ is a capabilty that we look into whether it is local or not.
+But even if $w$ is just an integer, we still need the capability we write through to have some kind of write permission.}
+
+After adding local capabilities, most of the definitions have changed slightly to take into the account new form the capabilties take.
+The only two instructions that have undergone enough change to deserve to be mentioned are $\store{}$ and $\restrict{}{}{}$.
+
+$\store{}$ has the same conditions as before, but if what it tries to write is a capability, then it needs to be a write access that permits write local.
+
+$\restrict{}{}{}$ looks like before, but now it uses a ``permission pair'' which is the access permission and whether it is local or not.
+Whether a $\restrict{}{}{}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above.
 
@@
 \newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
+\lau{Has not been adjusted since local capabilties were added.}
+
 Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\advb$.
@@
 \begin{align*}
-  \iota_{\var{start},\var{end}} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun \Regions\\
+  \iota_{\var{base},\var{end}} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun \Regions\\
   \iota_{\var{base},\var{end}} & \defeq \lambda \; \stdvr \ldotp ((\var{base},\var{end}),=,H_{\var{std}} (\stdvr)) \\ & \\
@@
 
+\subsection{Logical Relation}
+\lau{To be deleted when incorporated in the rest of the text.}
+\subsubsection{Worlds}
+\lau{disclaimer: the following is my current perception of what is going on}
+The worlds here are almost the same as the ones we defined in the case without local capabilities in Section~\ref{subsec:worlds}.
+The difference is that here our region keeps track of a view, namely whether it belongs to the \emph{permanent} or \emph{temporary} view of the current memory.
+The permanent part signifies what can be kept after a return to a caller whereas the temporary part will be lost after returning to a caller.
+
+Define $\Views$ to be
+\[
+  \Views \defeq \{\perma, \temp \}
+\]
+
+Assume a preordered \cofe{} $\Wor$ and an isomorphism $\xi$ such that:
+\[
+  \xi : \Wor \cong \blater (\nats \finparfun (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
+\]
+
+And the worlds are
+\[
+  \Worlds \defeq \nats \finparfun \Regions
+\]
+
 \section{Other examples and applications}

### 2016-10-26 capability-machine.tex
resolve conflict
Added missing notation in TD example
Added three small examples that could be interesting with local capabilities.

@@
 \label{sec:examples}
+\subsection{Simple Examples Using Local Capabilities}
+\label{subsec:example-loc-cap}
+\subsubsection{Simple use of local capabilities}
+\begin{verbatim}
+let f = fun adv =>
+          let l = 1 in
+          adv(l);
+          l := 1;
+          adv(0);
+          return *l
+\end{verbatim}
+In this example \texttt{let l = 1 in} allocates a new local capability \texttt{l} with read-write permissions.
+Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
+
+\subsection{Simple well-bracketness/local capability example}
+\begin{verbatim}
+let f = fun adv =>
+          let x = 1 in
+          adv();
+          assert(x == 1);
+          x := 2;
+          adv();
+          return 0
+\end{verbatim}
+A sensible calling scheme uses local capabilities for the continuation capability passed to the callee. In the above example, if \texttt{adv} can store the continuation, then 
+\texttt{adv} can invoke it during the second call and thus force the assertion to fail.
+
+\subsection{Another simple well-bracketness/local capability example}
+Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability.
+\begin{verbatim}
+let f = fun adv =>
+          let l = 0 in
+          let x = 1 in
+          adv(l);
+          assert(x == 1);
+          x := 2;
+          adv(0);
+          return 0
+\end{verbatim}
+In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}.
+\texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.
+
 \subsection{Ticket Dispenser}
@@
 \reta:     & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capability} \\
-           & \lea{r_0}{-2} & & \\
+           & \lea{\refreg{r_0}}{$-2$} & & \\
            & \load{\refreg{\rio}}{\refheap{r_0}} & & \\
 % retieve counter capability
-           & \lea{r_0}{1} & \bracket{2} & \annotate{2}{Retrieve the counter capability} \\
+           & \lea{\refreg{r_0}}{1} & \bracket{2} & \annotate{2}{Retrieve the counter capability} \\
            & \load{\refreg{r_1}}{\refheap{r_0}} & & \\

### 2016-10-27 capability-machine.tex
typos
Added suggestion for world and part of value relation

@@
 \newcommand{\lv}{\var{lv}}
-\newcommand{\gl}{g}
+\newcommand{\gl}{\var{g}}
 \newcommand{\pc}{\mathit{pc}}
@@
 \newcommand{\entryCond}[1]{\plainfun{entryCondition}{#1}}
+
+\newcommand{\revokeTemp}[1]{\plainfun{revokeTemp}{#1}}
+\newcommand{\erase}[2]{\floor*{#1}_{#2}}
 
@@
 
+%Views
+\newcommand{\plainview}[1]{\mathrm{#1}}
+\newcommand{\perma}{\plainview{perm}}
+\newcommand{\temp}{\plainview{temp}}
+\newcommand{\revoked}{\plainview{revoked}}
+
 %OP sem
@@
 \newcommand{\step}[1][]{\rightarrow_{#1}}
-
-%View
-\newcommand{\perma}{\mathrm{permanent}}
-\newcommand{\temp}{\mathrm{temporary}}
 
@@
 
+\section{Logical Relation'}
+\subsection{Worlds}
+Define the domain $\Views$ as follows
+\[
+  \Views = \{\perma,\temp,\revoked\}
+\]
+Using the ordering where $\revoked \future \temp$ and $\future$ is otherwise reflexive, and a family of equivalences where everything is equal at 0 and othervise the equivalences are ``normal'' equality.
+We claim this is a preordered \cofe{}.
+%TODO prove this claim!
+And show the new "region" is a preordered cofe with the claimed ordering.
+
+Assume $\Wor$ is a preordered \cofe{} and $\xi$ is an isomorphism such that
+\[
+  \xi : \Wor \cong \blater (\nats \finparfun (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
+\]
+The regions are here
+\[
+  \Regions = (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
+\]
+And the worlds are
+\[
+  \Worlds = \RegionNames \finparfun \Regions
+\]
+The ordering of worlds is
+
+ \begin{align*}
+ W' \future W & & \text{ iff } & &
+   \begin{gathered}
+     \dom(W') \supseteq \dom(W)\\ 
+     \text{ and } \\
+     \forall r \in \dom(W) \ldotp W'(r) \future W(r)
+   \end{gathered}
+ \end{align*}
+and the ordering of regions is
+\begin{gather*}
+  (v_2,s_2,\phi_2,H_2) \future (v_1,s_1,\phi_1,H_1) \\
+  \text{ iff } \\ 
+    \begin{gathered}
+      (v_2,v_1) = (\revoked,\revoked) \text{ and } (s_1,\phi_1,H_1) = (s_2,\phi_2,H_2) \\
+      \text{or} \\
+      (v_2,v_1) \in \{(\revoked,\temp), (\temp,\temp), (\perma,\perma) \} \text{ and } \\
+      (s_1,s_2) \in \phi_2 \text{ and } (\phi_1,H_1) = (\phi_2,H_2)
+    \end{gathered} 
+\end{gather*}
+The idea here is that the region is ``frozen'' after it has been revoked.
+
+Erase all but one view in a world:
+
+\begin{align*}
+  \erase{W}{v} \defeq \lambda r \ldotp 
+  \begin{cases}
+    W(r) & W(r).v = v\\
+    \undefined & \text{otherwise}
+  \end{cases}
+\end{align*}
+
+Standard region for when writing locally is permitted:
+\begin{align*}
+  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{pwl}}) \\ 
+  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+  \right\}
+\end{align*}
+
+Rewoking all temporary regions:
+
+\begin{align*}
+  \revokeTemp{} & : \Worlds \fun \Worlds \\
+  \revokeTemp{W} & \defeq \lambda r \ldotp 
+                   \begin{cases}
+                     (\revoked,s,\phi,H) & W(r) = (\temp,s,\phi,H) \\
+                     W(r)                & \text{otherwise}
+                   \end{cases}
+\end{align*}
+
+A standard region for when write local is not allowed:
+
+\begin{align*}
+  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{nwl}}) \\ 
+  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \\
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
+    \end{aligned}
+  \right\}
+\end{align*}
+
+\begin{align*}
+    & \writeCond{}(\iota,\var{regions})(\start,\addrend,W) =  \\
+    & \quad \begin{aligned}[t]
+              \{ n \mid & \;\exists r \in \var{regions} \ldotp \\
+              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+              & \;\qquad W(r)\nequal[n-1] \iota_{\start',\addrend'}(W(r).v) \}
+            \end{aligned}
+\end{align*}
+
+Now define the value relation as follows:
+\begin{align*}
+  \stdvr &\defeq \lambda \; V \ldotp 
+           \begin{aligned}[t]
+             & \{ \npair{i} \mid i \in \ints \} 
+             \union \\
+             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
+             \union \\
+             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \} 
+             \union \\
+             & \{ \npair{\stdcap[(\readwrite,\local)] } \mid \\
+             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(W))(\start,\addrend,W) \}
+             \union \\
+             & \{ \npair{\stdcap[(\readwrite,\glob)] } \mid  \\            
+             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
+             \union \\
+             & \{ \npair{\stdcap[(\readwritel,\local)] } \mid \\
+             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\var{regions})(\start,\addrend,W) \\
+             & \quad \text{for }\var{regions} = \dom(\erase{W}{\temp}) \union \dom(\erase{W}{\revoked})\}
+             \union \\
+             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \}
+             \union \\
+             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \} \union \\
+             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \} \union \\
+             & \{ \npair{\stdcap[(\rwlx,\local)]} \mid \}
+           \end{aligned}
+\end{align*}
+
 \section{Logical Relation}
@@
                        &  \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
-                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(\stdvr)(W') \ldotp \\
-                       & \quad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
+                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(\stdvr)(W') \ldotp \\
+                       & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
                  \end{aligned}
@@
 \begin{align*}
-  \stdvr & :  (\Worlds \monnefun \UPred{Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
+  \stdvr & :  (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
   \stdvr &\defeq \lambda \; V \ldotp \lambda \; W \ldotp 
@@
 \end{proof}
-
-\subsection{Logical Relation}
-\lau{To be deleted when incorporated in the rest of the text.}
-\subsubsection{Worlds}
-\lau{disclaimer: the following is my current perception of what is going on}
-The worlds here are almost the same as the ones we defined in the case without local capabilities in Section~\ref{subsec:worlds}.
-The difference is that here our region keeps track of a view, namely whether it belongs to the \emph{permanent} or \emph{temporary} view of the current memory.
-The permanent part signifies what can be kept after a return to a caller whereas the temporary part will be lost after returning to a caller.
-
-Define $\Views$ to be
-\[
-  \Views \defeq \{\perma, \temp \}
-\]
-
-Assume a preordered \cofe{} $\Wor$ and an isomorphism $\xi$ such that:
-\[
-  \xi : \Wor \cong \blater (\nats \finparfun (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
-\]
-
-And the worlds are
-\[
-  \Worlds \defeq \nats \finparfun \Regions
-\]
 

### 2016-10-28 capability-machine.tex
Spell checking

@@
 \documentclass{article}
+\usepackage[utf8]{inputenc}
 
@@
 \end{align*}
-Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capabaility is local or global.
+Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capability is local or global.
 Two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities.
@@
   large finite set.
-\item Table~\ref{tab:permission-list} describes what all the permissions grant acces to.
+\item Table~\ref{tab:permission-list} describes what all the permissions grant access to.
 \item Figure~\ref{fig:perm-hier} shows the ordering of the permissions, i.e, the elements of $\Perms$.
 \item Figure~\ref{fig:glob-hier} shows the ordering of $\local$ and $\glob$, i.e., the elements of $\Globals$.
-\item The ordering of $\Perms \times \Globals$ is pointwise.
+\item The ordering of $\Perms \times \Globals$ is point wise.
 \end{itemize}
@@
 The new condition depends on a case distinction on $w$.
-It is only if $w$ is a capabilty that we look into whether it is local or not.
+It is only if $w$ is a capability that we look into whether it is local or not.
 But even if $w$ is just an integer, we still need the capability we write through to have some kind of write permission.}
 
-After adding local capabilities, most of the definitions have changed slightly to take into the account new form the capabilties take.
+After adding local capabilities, most of the definitions have changed slightly to take into the account new form the capabilities take.
 The only two instructions that have undergone enough change to deserve to be mentioned are $\store{}$ and $\restrict{}{}{}$.
@@
 \newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
-\lau{Has not been adjusted since local capabilties were added.}
+\lau{Has not been adjusted since local capabilities were added.}
 
@@
 % Set io to -1
-\initb:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialize io to $-1$} \\
+\initb:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialise io to $-1$} \\
 % Store io capability just after jump to adv
@@
 \]
-Using the ordering where $\revoked \future \temp$ and $\future$ is otherwise reflexive, and a family of equivalences where everything is equal at 0 and othervise the equivalences are ``normal'' equality.
+Using the ordering where $\revoked \future \temp$ and $\future$ is otherwise reflexive, and a family of equivalences where everything is equal at 0 and otherwise the equivalences are ``normal'' equality.
 We claim this is a preordered \cofe{}.
@@
 
-Rewoking all temporary regions:
+Revoking all temporary regions:
 
@@
 
-This cannot be solved with sets, so we use preordered complete ordered families of equivalences where it is possible to solve such an equation that ressembles the above one, namely it is possible to find an isomorphism $\xi$ and preordered \cofe{} $W$ such that
+This cannot be solved with sets, so we use preordered complete ordered families of equivalences where it is possible to solve such an equation that resembles the above one, namely it is possible to find an isomorphism $\xi$ and preordered \cofe{} $W$ such that
 \[
@@
   An ordered family of equivalences (o.f.e.) is a set and a family of equivalences, $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$.
-The familly of equivalences have to satisfy the following properties
+The family of equivalences have to satisfy the following properties
   \begin{itemize}
@@
 \begin{definition}[\cofes]
-  A complete orderede family of equivalences is an o.f.e.\ $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ where all Cauchy sequences in $X$ have a limit in $X$.
+  A complete ordered family of equivalences is an o.f.e.\ $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ where all Cauchy sequences in $X$ have a limit in $X$.
 \end{definition}
@@
 \begin{definition}[Preordered \cofes]
-  A preordered \cofe{}\ is a \cofe{}\ equiped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
+  A preordered \cofe{}\ is a \cofe{}\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
   \begin{itemize}
@@
 
-The category of \cofes{} is the category whith \cofes{} as objects and non-expansive functions as morphisms.
+The category of \cofes{} is the category with \cofes{} as objects and non-expansive functions as morphisms.
 We denote this category $\CatC$.
@@
 \end{lemma}
-With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-copmosition of $\blater$:
+With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-composition of $\blater$:
 \[
@@
 \]
-We have a theorem that gives us a solution to the recurisve domain equation
+We have a theorem that gives us a solution to the recursive domain equation
 \[
@@
 \]
-The solution to the recursice domain equations is presented by \cite{Birkedal:2010:TCS:411:4102-4122}.
+The solution to the recursive domain equations is presented by \cite{Birkedal:2010:TCS:411:4102-4122}.
 They solve it in pre-ordered, non-empty, complete, 1-bounded ultrametric spaces, but they have a simple correspondence to pre-ordered \cofes{}.
@@
 Our logical relation is defined using multiple recursive definitions, so the definitions in the following subsections are defined simultaneously.
-We want to define the value relation as the fixed-point given by Banach's fixed-point theorem, so all our definitions will be parameterized with the value relation.
+We want to define the value relation as the fixed-point given by Banach's fixed-point theorem, so all our definitions will be parameterised with the value relation.
 \subsubsection{Observation Relation}
@@
 
-\subsubsection{Regiser-File Relation}
+\subsubsection{Register-File Relation}
 \label{subsubsec:register-file-relation}
@@
 \label{subsubsec:capability-conditions}
-The definition of the value relation has the same conditions several times, so to define it consisely, we define the following conditions.
+The definition of the value relation has the same conditions several times, so to define it concisely, we define the following conditions.
 \begin{align*}
@@
 \lau{This section introduces the definitions, we started working with, so it is here so we don't have to rewrite everything.}
-Here we state the definitions, we will use for pratical purposes.
+Here we state the definitions, we will use for practical purposes.
 Assume $V$ is the fixed point given by Banach's fixed point theorem as described in Section~\ref{subsubsec:value-relation}.
@@
 
-For natural number $n$, adresses $\start$, $\addrend$, and $\addr$, permission $\perm$, and world $W$, we define the capability conditions as follows
+For natural number $n$, addresses $\start$, $\addrend$, and $\addr$, permission $\perm$, and world $W$, we define the capability conditions as follows
 \begin{align*}
@@
 See Section~\ref{subsubsec:standard-region}.
-We here define another usefull region, namely the one that requires all the words in the region to be the constant 0.
+We here define another useful region, namely the one that requires all the words in the region to be the constant 0.
 \[
@@
   \item $H_{\var{std}}^0$ is non-expansive.
-  \item For all adresses $\start$
+  \item For all addresses $\start$
     and $\addrend$, $ H_{\var{std}}^0 \; (\start,\addrend)$ is monotone and non-expansive.
@@
 If $r \neq \pcreg$, then if $r= r_0$, then by assumption \ref{eq:ftlr:cont-in-stdkr}, we have $\npair[n']{c} \in \stdkr(W)$ which entails $\npair[n']{c} \in \stdvr(W)$ which together with uniformity gives the $\writeCond{}$ due to the permission.
-Otherwise, the same folloes from assumption \ref{eq:ftlr:reg-in-stdrr} and uniformity.
+Otherwise, the same follows from assumption \ref{eq:ftlr:reg-in-stdrr} and uniformity.
 
@@
 return of any adversary function, but this could cause an important overhead.
-Perhaps the processor should accomodate this with a special instruction that can
+Perhaps the processor should accommodate this with a special instruction that can
 zero the entire array that a capability points to?
@@
 \begin{definition}[Set \cofe{}\ Construction]
-  Given a set $X$, then $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ is a \cofe{}\ where all the equivalences are equiality.
+  Given a set $X$, then $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ is a \cofe{}\ where all the equivalences are equality.
 \end{definition}
@@
   \]
-  let $n$ be givenm show
+  let $n$ be given show
   \[
@@
 \begin{proof}[Proof of Lemma~\ref{lem:R-func}]
-   $R(f) : R(X) \rightarrow R(Y)$ is \emph{monotone} for $f : X \nefun Y$. Given aformentioned $f$. Further let $(s_2,\Phi_2,H_2) \future (s_1,\Phi_1,H_1)$ be given.
+   $R(f) : R(X) \rightarrow R(Y)$ is \emph{monotone} for $f : X \nefun Y$. Given aforementioned $f$. Further let $(s_2,\Phi_2,H_2) \future (s_1,\Phi_1,H_1)$ be given.
 
@@
 
-  Idenntity distributes over $R$, $R(\id) = \id$.
+  Identity distributes over $R$, $R(\id) = \id$.
   \[
@@
 \]
-which is what we neede to show.
+which is what we needed to show.
 
@@
 First we show $W_1' \nequal W_2'$.
-The domains are defined to be equal, so that is satsfied.
+The domains are defined to be equal, so that is satisfied.
 Let $r \in \dom(W_1')$ and show $W_1' \; r \nequal W_2' \; r$.
@@
 Here the two last are trivial.
-The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notive that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
+The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notice that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
 \end{proof}

### 2016-10-28 capability-machine.tex
comments inserted about examples

@@
 \newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
-\newcommand\kars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+\newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
 
@@
 Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
+\lars{So this means that the value returned by *l is 1, correct ? We
+  could add an assert statement to the example to show it more directly.}
+
+\lars{What about an even simpler example, such as the one below:}
+\begin{verbatim}
+let f = fun adv =>
+          let l = 1 in
+          adv();
+          assert (l == 1)
+\end{verbatim}
 
@@
 \texttt{adv} can invoke it during the second call and thus force the assertion to fail.
+\lars{don't you want to allocate the location \texttt{x} first and then
+  return a function taking \texttt{adv} as an argument ??}
 

### 2016-10-28 capability-machine.tex
comments

@@
           adv(0);
-          return *l
+          assert(l == 1)
 \end{verbatim}
@@
   could add an assert statement to the example to show it more directly.}
+\lau{Yes, and done.}
 
@@
 \end{verbatim}
-
-\subsection{Simple well-bracketness/local capability example}
+\lau{This example should certainly also hold true work, but it does not need local capabilities (it might serve well as a sanity check). }
+
+\subsection{Simple well-bracketedness/local capability example}
 \begin{verbatim}
@@
   return a function taking \texttt{adv} as an argument ??}
-
-\subsection{Another simple well-bracketness/local capability example}
+\lau{I am not sure I get what you have in mind, could you perhaps write it?}
+
+\subsection{Another simple well-bracketedness/local capability example}
 Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability.
@@
   &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{pwl}}) \\ 
-  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
+  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
   &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{nwl}}) \\ 
-  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
+  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
 \begin{align*}
-  \stdvr &\defeq \lambda \; V \ldotp 
+  \stdvr &\defeq \lambda \; W \ldotp 
            \begin{aligned}[t]
@@
              & \{ \npair{\stdcap[(\readwrite,\glob)] } \mid  \\            
-             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
+             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
              \union \\

### 2016-10-31 capability-machine.tex
LR suggestion

@@
 \newcommand{\parfun}{\rightharpoonup}
-\newcommand{\finparfun}{\stackrel{\textit{\tiny{fin}}}{\rightharpoonup}}
-\newcommand{\monnefun}{\stackrel{\textit{\tiny{mon, ne}}}{\rightarrow}}
-\newcommand{\monfun}{\stackrel{\textit{\tiny{mon}}}{\rightarrow}}
-\newcommand{\nefun}{\stackrel{\textit{\tiny{ne}}}{\rightarrow}}
+\newcommand{\finparfun}{\xrightharpoonup{\textit{\tiny{fin}}}}
+\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
+\newcommand{\monfun}{\xrightarrow{\textit{\tiny{mon}}}}
+\newcommand{\nefun}{\xrightarrow{\textit{\tiny{ne}}}}
 \newcommand{\fun}{\rightarrow}
@@
 \newcommand{\revokeTemp}[1]{\plainfun{revokeTemp}{#1}}
-\newcommand{\erase}[2]{\floor*{#1}_{#2}}
+\newcommand{\erase}[2]{\floor*{#1}_{\{#2\}}}
 
@@
 \newcommand{\future}{\mathbin{\sqsupseteq}}
+\newcommand{\futurewk}{\mathbin{\sqsupseteq}_{\var{wk}}}
+\newcommand{\futurestr}{\mathbin{\sqsupseteq}_{\var{str}}}
 \newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
+
+\newcommand{\monwknefun}{\xrightarrow[\text{\tiny{$\futurewk$}}]{\textit{\tiny{mon, ne}}}}
+\newcommand{\monstrnefun}{\xrightarrow[\text{\tiny{$\futurestr$}}]{\textit{\tiny{mon, ne}}}}
 
@@
 \newcommand{\Regions}{\plaindom{Region}}
+\newcommand{\Reg}{\plaindom{Reg}}
+\newcommand{\Regwk}{\Reg_{\futurewk}}
+\newcommand{\Regstr}{\Reg_{\futurestr}}
 \newcommand{\Worlds}{\plaindom{World}}
@@
 \newcommand{\glob}{\plainperm{global}}
+
+\newcommand{\localReg}{\var{localReg}}
+\newcommand{\globalReg}{\var{globalReg}}
 
@@
 \label{subsec:example-loc-cap}
+In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate with the return value 1.
+Notice that an assertion will never terminate in a $\failed$ state.
 \subsubsection{Simple use of local capabilities}
@@
 \texttt{adv} can invoke it during the second call and thus force the assertion to fail.
+We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
 \lars{don't you want to allocate the location \texttt{x} first and then
   return a function taking \texttt{adv} as an argument ??}
-\lau{I am not sure I get what you have in mind, could you perhaps write it?}
+\lau{I am not sure I get what you have in mind, could you perhaps write the program you had in mind?}
 
@@
 \subsection{Worlds}
-Define the domain $\Views$ as follows
-\[
-  \Views = \{\perma,\temp,\revoked\}
-\]
-Using the ordering where $\revoked \future \temp$ and $\future$ is otherwise reflexive, and a family of equivalences where everything is equal at 0 and otherwise the equivalences are ``normal'' equality.
-We claim this is a preordered \cofe{}.
-%TODO prove this claim!
-And show the new "region" is a preordered cofe with the claimed ordering.
-
-Assume $\Wor$ is a preordered \cofe{} and $\xi$ is an isomorphism such that
-\[
-  \xi : \Wor \cong \blater (\nats \finparfun (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
-\]
-The regions are here
-\[
-  \Regions = (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
-\]
+Assume $\Regstr$ and $\Regwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
+\begin{align*}
+  \xi_{\var{wk}} :   \Regwk \cong \blater (&\{\revoked\} \union \\
+                                           & \{\temp\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})) \union \\
+                                           & \{\perma\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}} \\ \\
+  \xi_{\var{str}} :   \Regstr \cong \blater (&\{\revoked\} \union \\
+                                           & \{\temp\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})) \union \\
+                                           & \{\perma\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}}
+\end{align*}
+We now define the regions to be
+\begin{align*}
+  \Regions = & \{\revoked\} \union \\
+             & \{\temp\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})) \union \\
+             & \{\perma\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments}))
+\end{align*}
 And the worlds are
@@
 \]
-The ordering of worlds is
-
- \begin{align*}
- W' \future W & & \text{ iff } & &
+Where $\RegionNames = \nats$.
+
+Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system:
+\begin{gather*}
+  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
+  \text{ iff } \\ 
+  (s_1,s_2) \in \phi_2 \text{ and } (\phi_1,H_1) = (\phi_2,H_2)
+\end{gather*}
+our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
+\begin{align*}
+  &\quad r_2 \futurewk r_1 \\
+  &\text{ iff } \\ 
+  &\quad\exists v_2,s_2,\phi_2,H_2,v_1,s_1,\phi_1,H_1\ldotp\\
+  &\qquad      r_2 = (v_2,s_2,\phi_2,H_2) \text{ and } \\
+  &\qquad r_1=(v_1,s_1,\phi_1,H_1) \text{ and }\\
+  &\qquad v_1 = v_2 \text{ and } (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1)\\
+  &\quad \text{or} \\ 
+  &\qquad r_1 = r_2 = \revoked
+\end{align*}
+For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
+\begin{align*}
+  &\quad r_2 \futurestr r_1 \\
+  &\text{ iff } \\ 
+  &\quad r_2 \futurewk r_1 \\
+  &\quad\text{or} \\ 
+  &\quad\exists s_1,\phi_1,H_1\ldotp\\
+  &\qquad r_2 = \revoked \text{ and } r_1 = (\temp,s_1,\phi_1,H_1)
+\end{align*}
+
+We now get to future world relations $\futurewk$ and $\futurestr$ are defined as one would expect.
+That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
+ \begin{align*}
+ W' \future_{\var{wk}/\var{str}} W & & \text{ iff } & &
    \begin{gathered}
@@
      \text{ and } \\
-     \forall r \in \dom(W) \ldotp W'(r) \future W(r)
+     \forall r \in \dom(W) \ldotp W'(r) \future_{\var{wk}/\var{str}} W(r)
    \end{gathered}
  \end{align*}
-and the ordering of regions is
-\begin{gather*}
-  (v_2,s_2,\phi_2,H_2) \future (v_1,s_1,\phi_1,H_1) \\
-  \text{ iff } \\ 
-    \begin{gathered}
-      (v_2,v_1) = (\revoked,\revoked) \text{ and } (s_1,\phi_1,H_1) = (s_2,\phi_2,H_2) \\
-      \text{or} \\
-      (v_2,v_1) \in \{(\revoked,\temp), (\temp,\temp), (\perma,\perma) \} \text{ and } \\
-      (s_1,s_2) \in \phi_2 \text{ and } (\phi_1,H_1) = (\phi_2,H_2)
-    \end{gathered} 
-\end{gather*}
-The idea here is that the region is ``frozen'' after it has been revoked.
-
-Erase all but one view in a world:
-
-\begin{align*}
-  \erase{W}{v} \defeq \lambda r \ldotp 
+
+Erase all but a set of views.
+
+\begin{align*}
+  \erase{W}{S} \defeq \lambda r \ldotp 
   \begin{cases}
-    W(r) & W(r).v = v\\
+    W(r) & W(r).v \in S\\
     \undefined & \text{otherwise}
@@
 
-Standard region for when writing locally is permitted:
-\begin{align*}
-  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{pwl}}) \\ 
-  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+Standard regions for when writing locally is permitted:
+\begin{align*}
+  &\iota_{\start,\addrend}^{\var{pwl}} : \Regions\\
+  &\iota_{\start,\addrend}^{\var{pwl}} \defeq (\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
+  &H^{\var{pwl}} : \States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{pwl}} \; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+      &\exists \start,\addrend \ldotp s = (\start,\addrend) \land \\
+      &\quad \dom(\hs) = [\start,\addrend] \land \\
+      &\quad \forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\hat{W})
     \end{aligned}
@@
 \end{align*}
-
 Revoking all temporary regions:
-
 \begin{align*}
@@
                    \begin{cases}
-                     (\revoked,s,\phi,H) & W(r) = (\temp,s,\phi,H) \\
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
                      W(r)                & \text{otherwise}
@@
 \end{align*}
-
-A standard region for when write local is not allowed:
-
-\begin{align*}
-  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{nwl}}) \\ 
-  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+Standard regions for when write local is not allowed:
+\begin{align*}
+  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
+  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
+    \begin{cases}
+      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
+      \bot & \text{otherwise}
+\end{cases}
+\\ \\
+  &H^{\var{nwl}} : \States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{nwl}} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
+      &\exists \start,\addrend \ldotp s = (\start,\addrend) \land \\
       &\dom(\hs) = [\start,\addrend] \land \\
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\hat{W}}) %Does it make sense to use revokeTemp here when it does not "match up" with the world?
     \end{aligned}
@@
 
-\begin{align*}
-    & \writeCond{}(\iota,\var{regions})(\start,\addrend,W) =  \\
+For convenience define
+\begin{align*}
+  \localReg & \defeq \dom(\erase{W}{\perm,\temp}) \\
+  \globalReg & \defeq \dom(\erase{W}{\perm})
+\end{align*}
+$\localReg$ are the regions that local capabilities may govern. $\globalReg$ are the regions that global capabilities may govern. Now define the following function
+\[
+  \var{localityReg}(\gl) \defeq 
+  \begin{cases}
+    \localReg & \gl = \local \\
+    \globalReg & \gl = \glob
+  \end{cases}
+\]
+
+\begin{align*}
+    & \writeCond{}(\iota,\gl)(\start,\addrend,W) =  \\
     & \quad \begin{aligned}[t]
-              \{ n \mid & \;\exists r \in \var{regions} \ldotp \\
+              \{ n \mid & \;\exists r \in \var{localityReg}(g) \ldotp \\
               & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
@@
 
+\begin{align*}
+    & \readCond{}(\gl)(\start,\addrend,W) =  \\
+    & \quad \begin{aligned}[t]
+              \{ n \mid & \;\exists r \in \var{localityReg}(g) \ldotp \\
+              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+              & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
+            \end{aligned}
+\end{align*}
+\lau{I am in doubt whether it should be $\iota^{\var{pwl}}$ in the read condition.}
+
+\begin{align*}
+  & \execCond{}(\gl)(\perm,\start,\addrend,W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ n \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad\forall a \in [\start,\addrend] \ldotp\\
+      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+\begin{align*}
+  & \entryCond{}(\gl)(\start,\addrend,\addr,W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ n \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
 Now define the value relation as follows:
 \begin{align*}
-  \stdvr &\defeq \lambda \; W \ldotp 
+  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
+  &\stdvr\defeq \lambda \; W \ldotp 
            \begin{aligned}[t]
@@
              \union \\
-             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \} 
+             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
+             & \quad n \in \readCond{}(\gl)(\start,\addrend,W)\} 
              \union \\
-             & \{ \npair{\stdcap[(\readwrite,\local)] } \mid \\
-             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(W))(\start,\addrend,W) \}
+             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
+             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\gl)(\start,\addrend,W) \}
              \union \\
-             & \{ \npair{\stdcap[(\readwrite,\glob)] } \mid  \\            
-             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
+             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
+             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\gl)(\start,\addrend,W) \} \union \\
+             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
+             & \quad n \in \readCond{}(\gl)(\start,\addrend,W) \land \\
+             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
              \union \\
-             & \{ \npair{\stdcap[(\readwritel,\local)] } \mid \\
-             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\var{regions})(\start,\addrend,W) \\
-             & \quad \text{for }\var{regions} = \dom(\erase{W}{\temp}) \union \dom(\erase{W}{\revoked})\}
+             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
+             & \quad n \in \entryCond{}(\gl)(\start,\addrend,\addr,W)\} 
              \union \\
-             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \}
+             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
+             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\gl)(\start,\addrend,W) \land\\
+             & \quad n \in \execCond{}(\gl)(\rwx,\start,\addrend,W)  \land \\
+             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
              \union \\
-             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \} \union \\
-             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \} \union \\
-             & \{ \npair{\stdcap[(\rwlx,\local)]} \mid \}
+             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
+             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\gl)(\start,\addrend,W) \land\\
+             & \quad n \in \execCond{}(\gl)(\rwlx,\start,\addrend,W) \land \\
+             & \quad n \in \execCond{}(\gl)(\rwx,\start,\addrend,W) \land \\
+             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
+           \end{aligned}
+\end{align*}
+
+\begin{align*}
+  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
+  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
+                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                           & \qquad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
+\end{align*}
+
+\begin{align*}
+  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
+  \stdrr \defeq & \lambda W \ldotp
+                     \begin{aligned}[t]
+                       \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                                            & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
+                     \end{aligned}
+\end{align*}
+
+\begin{align*}
+  \stdkr : &  \Worlds \monwknefun \UPred{\Words} \\
+  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
+                     \begin{aligned}[t]
+                       & \npair{c} \in \stdvr(W) \land \\
+                       &  \forall W' \futurewk W, n' < n\ldotp
+                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(\stdvr)(W') \ldotp \\
+                       & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
+                 \end{aligned}
+\end{align*}
+
+\begin{align*}
+  \stder : & \Worlds \nefun \UPred{\Words} \\
+  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
+  \begin{aligned}[t]
+    & \forall n' \leq n \ldotp\\
+    & \quad \forall \npair[n']{\reg} \in \stdrr(\stdvr)(W) \ldotp \\
+    & \quad \quad \forall \npair[n']{c} \in \stdkr(\stdvr)(W) \ldotp \\
+    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
+    & \quad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
   \end{aligned}
@@
     \text{ iff } \\
-    \forall s \in \dom(f) \ldotp f(s) \future g(s)
+    \forall s \in \dom(g) \ldotp f(s) \future g(s)
   \end{gathered}
@@
     \text{ iff } \\
-    \forall s \in \dom(f) \ldotp f(s) \future g(s)
+    \forall s \in \dom(g) \ldotp f(s) \future g(s) 
   \end{gathered}

### 2016-11-03 capability-machine.tex
.

@@
     W(r) & W(r).v \in S\\
-    \undefined & \text{otherwise}
+    \bot & \text{otherwise}
   \end{cases}
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^{\var{pwl}} : \Regions\\
-  &\iota_{\start,\addrend}^{\var{pwl}} \defeq (\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
+  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
+  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
   &H^{\var{pwl}} : \States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})\\
@@
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\hat{W}}) %Does it make sense to use revokeTemp here when it does not "match up" with the world?
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\hat{W}})
     \end{aligned}
@@
 \end{align*}
+\lau{In the above $H$ definitions, $\hat{W}$ is not a world, so there is a type problem that needs to be fixed. Maybe solve the recursive domain equations wrt the worlds instead.}
 
@@
 \begin{align*}
+    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats}  \\
     & \writeCond{}(\iota,\gl)(\start,\addrend,W) =  \\

### 2016-11-03 capability-machine.tex
Reverted change with general state in H definitions+ typo

@@
 
-We now get to future world relations $\futurewk$ and $\futurestr$ are defined as one would expect.
+We now get two future world relations $\futurewk$ and $\futurestr$ are defined as one would expect.
 That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
@@
 
-Erase all but a set of views.
-
+Erase all but a set of views:
 \begin{align*}
@@
   &H^{\var{pwl}} : \States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{pwl}} \; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
-      &\exists \start,\addrend \ldotp s = (\start,\addrend) \land \\
-      &\quad \dom(\hs) = [\start,\addrend] \land \\
-      &\quad \forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\hat{W})
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\hat{W})
     \end{aligned}
@@
     \begin{aligned}
-      &\exists \start,\addrend \ldotp s = (\start,\addrend) \land \\
       &\dom(\hs) = [\start,\addrend] \land \\

### 2016-11-03 capability-machine.tex
Changed the recursive domain equation to be recursive in the worlds

@@
 \newcommand{\Reg}{\plaindom{Reg}}
-\newcommand{\Regwk}{\Reg_{\futurewk}}
-\newcommand{\Regstr}{\Reg_{\futurestr}}
 \newcommand{\Worlds}{\plaindom{World}}
 \newcommand{\Wor}{\plaindom{Wor}}
+\newcommand{\Worwk}{\Wor_{\futurewk}}
+\newcommand{\Worstr}{\Wor_{\futurestr}}
+\newcommand{\xiwk}{\xi_{\var{wk}}}
+\newcommand{\xistr}{\xi_{\var{str}}}
 \newcommand{\StorePred}{\plaindom{HeapSegPred}}
@@
 \subsection{Worlds}
-Assume $\Regstr$ and $\Regwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
-\begin{align*}
-  \xi_{\var{wk}} :   \Regwk \cong \blater (&\{\revoked\} \union \\
-                                           & \{\temp\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})) \union \\
-                                           & \{\perma\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}} \\ \\
-  \xi_{\var{str}} :   \Regstr \cong \blater (&\{\revoked\} \union \\
-                                           & \{\temp\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})) \union \\
-                                           & \{\perma\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}}
+Assume $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
+\begin{align*}
+  \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (&\{\revoked\} \union \\
+                                           & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \union \\
+                                           & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}}) \\ \\
+  \xi_{\var{str}} :   \Worstr \cong (\nats \finparfun \blater (&\{\revoked\} \union \\
+                                           & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \union \\
+                                           & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}})
 \end{align*}
@@
   \Regions = & \{\revoked\} \union \\
-             & \{\temp\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})) \union \\
-             & \{\perma\} \times \States \times \Rel \times (\States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments}))
+             & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \union \\
+             & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))
 \end{align*}
@@
   &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
-  &H^{\var{pwl}} : \States \fun ((\nats \finparfun \Regwk) \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
   &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
@@
       &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\hat{W})
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
     \end{aligned}
@@
 \\ \\
-  &H^{\var{nwl}} : \States \fun ((\nats \finparfun \Regstr) \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
   &H^{\var{nwl}} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
@@
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\hat{W}})
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
     \end{aligned}

### 2016-11-03 capability-machine.tex
Typos

@@
-\documentclass{article}
+\documentclass[a4paper]{article}
+%\usepackage[a4paper]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \newcommand{\forcenewline}{$\phantom{v}$\\}
+\newcommand{\judgment}[2]{\paragraph{#1}\hspace{\stretch{1}}\fbox{$#2$}}
 
@@
     \node[main node] (2) [below right of=1] {$\exec$};
-    \node[main node] (3) [below of=2] {$\entry$};
+    \node[main node] (3) [below right of=2] {$\entry$};
     \node[main node] (4) [below left of=1] {$\readwrite$};
-    \node[main node] (5) [below of=4] {$\readonly$};
+    \node[main node] (5) [below right of=4] {$\readonly$};
     \node[main node] (6) [below right of=5] {$\noperm$};
@@
     (2) edge (3)
+    (2) edge (5)
     (3) edge (6)
@@
 \begin{align*}
-  \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (&\{\revoked\} \union \\
-                                           & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \union \\
+  \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
+                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
                                                                 & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}}) \\ \\
-  \xi_{\var{str}} :   \Worstr \cong (\nats \finparfun \blater (&\{\revoked\} \union \\
-                                           & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \union \\
+  \xi_{\var{str}} :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
+                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
                                                                 & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}})
@@
 \begin{align*}
-  \Regions = & \{\revoked\} \union \\
-             & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \union \\
+  \Regions = & \{\revoked\} \uplus \\
+             & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \uplus \\
              & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))
@@
   &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{nwl}} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
 \begin{align*}
-  \localReg & \defeq \dom(\erase{W}{\perm,\temp}) \\
-  \globalReg & \defeq \dom(\erase{W}{\perm})
-\end{align*}
-$\localReg$ are the regions that local capabilities may govern. $\globalReg$ are the regions that global capabilities may govern. Now define the following function
-\[
-  \var{localityReg}(\gl) \defeq 
+  \localReg(W) & \defeq \dom(\erase{W}{\perm,\temp}) \\
+  \globalReg(W) & \defeq \dom(\erase{W}{\perm})
+\end{align*}
+$\localReg$ are the regions that local capabilities may govern - that is permanent and global regions.
+$\globalReg$ are the regions that global capabilities may govern - that is permanent regions.
+Now define the following function
+\[
+  \var{localityReg}(\gl,W) \defeq 
   \begin{cases}
-    \localReg & \gl = \local \\
-    \globalReg & \gl = \glob
+    \localReg(W) & \text{if } \gl = \local \\
+    \globalReg(W) & \text{if } \gl = \glob
   \end{cases}
@@
     & \quad \begin{aligned}[t]
-              \{ n \mid & \;\exists r \in \var{localityReg}(g) \ldotp \\
+              \{ n \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
               & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
@@
   \stdkr : &  \Worlds \monwknefun \UPred{\Words} \\
-  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
+  \stdkr \defeq & \lambda W \ldotp\{ \npair{c} \mid
                      \begin{aligned}[t]
@@
                        &  \forall W' \futurewk W, n' < n\ldotp
-                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(\stdvr)(W') \ldotp \\
+                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(W') \ldotp \\
                        & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
@@
     & \forall n' \leq n \ldotp\\
-    & \quad \forall \npair[n']{\reg} \in \stdrr(\stdvr)(W) \ldotp \\
-    & \quad \quad \forall \npair[n']{c} \in \stdkr(\stdvr)(W) \ldotp \\
+    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
+    & \quad \quad \forall \npair[n']{c} \in \stdkr(W) \ldotp \\
     & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\

### 2016-11-03 capability-machine.tex
Removed continuation relation! Expression relation now uses value relation instead (indirectly throug ref-file rel).
Changed the read write condition into a pure write condition.
Changed the conditions to use UPreds for readability. Changed some definitions to inference rules for readability.

@@
 \usepackage{verbatim}
+\usepackage{mathpartir}
 
@@
 \newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
-\newcommand{\writeCond}[1]{\plainfun{readWriteCondition}{#1}}
+\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
 \newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
@@
 Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system:
-\begin{gather*}
-  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
-  \text{ iff } \\ 
-  (s_1,s_2) \in \phi_2 \text{ and } (\phi_1,H_1) = (\phi_2,H_2)
-\end{gather*}
+\begin{mathpar}
+  \inferrule{  (s_1,s_2) \in \phi_2 \\
+               (\phi_1,H_1) = (\phi_2,H_2)}
+            {  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) }
+\end{mathpar}
 our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
-\begin{align*}
-  &\quad r_2 \futurewk r_1 \\
-  &\text{ iff } \\ 
-  &\quad\exists v_2,s_2,\phi_2,H_2,v_1,s_1,\phi_1,H_1\ldotp\\
-  &\qquad      r_2 = (v_2,s_2,\phi_2,H_2) \text{ and } \\
-  &\qquad r_1=(v_1,s_1,\phi_1,H_1) \text{ and }\\
-  &\qquad v_1 = v_2 \text{ and } (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1)\\
-  &\quad \text{or} \\ 
-  &\qquad r_1 = r_2 = \revoked
-\end{align*}
+\begin{mathpar}
+  \inferrule{ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
+               v_1 = v_2 }
+            { (v_2,s_2,\phi_2,H_2) \futurewk (v_1,s_1,\phi_1,H_1) }
+\and
+  \inferrule{ }
+            { \revoked \futurewk \revoked }
+\end{mathpar}
 For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
-\begin{align*}
-  &\quad r_2 \futurestr r_1 \\
-  &\text{ iff } \\ 
-  &\quad r_2 \futurewk r_1 \\
-  &\quad\text{or} \\ 
-  &\quad\exists s_1,\phi_1,H_1\ldotp\\
-  &\qquad r_2 = \revoked \text{ and } r_1 = (\temp,s_1,\phi_1,H_1)
-\end{align*}
+\begin{mathpar}
+  \inferrule{ r_2 \futurewk  r_1 }
+            { r_2 \futurestr r_1 }
+\and
+  \inferrule{ }
+            { \revoked \futurestr (\temp,s,\phi,H)}
+\end{mathpar}
 
@@
 That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
- \begin{align*}
- W' \future_{\var{wk}/\var{str}} W & & \text{ iff } & &
-   \begin{gathered}
-     \dom(W') \supseteq \dom(W)\\ 
-     \text{ and } \\
-     \forall r \in \dom(W) \ldotp W'(r) \future_{\var{wk}/\var{str}} W(r)
-   \end{gathered}
- \end{align*}
+\begin{mathpar}
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+              \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
+            { W' \futurestr W }
+\and
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+              \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
+            { W' \futurestr W }
+\end{mathpar}
 
@@
 \begin{align*}
-    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats}  \\
-    & \writeCond{}(\iota,\gl)(\start,\addrend,W) =  \\
+    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+    & \writeCond{}(\iota,\gl)(W) =  \\
     & \quad \begin{aligned}[t]
-              \{ n \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
               & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-              & \;\qquad W(r)\nequal[n-1] \iota_{\start',\addrend'}(W(r).v) \}
+              & \;\qquad W(r)\nsupeq[n-1] \iota_{\start',\addrend'}(W(r).v) \}
             \end{aligned}
@@
 \begin{align*}
-    & \readCond{}(\gl)(\start,\addrend,W) =  \\
+    & \readCond{}(\gl)(W) =  \\
     & \quad \begin{aligned}[t]
-              \{ n \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
                         & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
@@
 \begin{align*}
-  & \execCond{}(\gl)(\perm,\start,\addrend,W) = \\
+  & \execCond{}(\gl)(W) = \\
   & \quad
     \begin{aligned}[t]
-      \{ n \mid &  \forall n' < n \ldotp\\
+      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
       & \quad \forall W' \future W \ldotp\\
@@
 \begin{align*}
-  & \entryCond{}(\gl)(\start,\addrend,\addr,W) = \\
+  & \entryCond{}(\gl)(W) = \\
   & \quad
     \begin{aligned}[t]
-      \{ n \mid &  \forall n' < n \ldotp\\
+      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
       & \quad \forall W' \future W \ldotp\\
@@
              & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
-             & \quad n \in \readCond{}(\gl)(\start,\addrend,W)\} 
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
              \union \\
              & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
-             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\gl)(\start,\addrend,W) \}
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \}
              \union \\
              & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
-             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\gl)(\start,\addrend,W) \} \union \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \}
+             \union \\
              & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
-             & \quad n \in \readCond{}(\gl)(\start,\addrend,W) \land \\
-             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
              \union \\
              & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
-             & \quad n \in \entryCond{}(\gl)(\start,\addrend,\addr,W)\} 
+             & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
              \union \\
              & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
-             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\gl)(\start,\addrend,W) \land\\
-             & \quad n \in \execCond{}(\gl)(\rwx,\start,\addrend,W)  \land \\
-             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \land\\
+             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
+             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
              \union \\
              & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
-             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\gl)(\start,\addrend,W) \land\\
-             & \quad n \in \execCond{}(\gl)(\rwlx,\start,\addrend,W) \land \\
-             & \quad n \in \execCond{}(\gl)(\rwx,\start,\addrend,W) \land \\
-             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \land\\
+             & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
+             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
+             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
            \end{aligned}
@@
   \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
-                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
-                           & \qquad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
+  \observations (W) \defeq & \lambda W \ldotp 
+                             \begin{aligned}[t]
+                               \{ \npair{(\reg,\hs)} \mid & \forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                                                          & \quad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'} \}
+                             \end{aligned}
 \end{align*}
@@
 \begin{align*}
-  \stdkr : &  \Worlds \monwknefun \UPred{\Words} \\
-  \stdkr \defeq & \lambda W \ldotp\{ \npair{c} \mid
-                     \begin{aligned}[t]
-                       & \npair{c} \in \stdvr(W) \land \\
-                       &  \forall W' \futurewk W, n' < n\ldotp
-                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(W') \ldotp \\
-                       & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
-                 \end{aligned}
-\end{align*}
-
-\begin{align*}
   \stder : & \Worlds \nefun \UPred{\Words} \\
@@
     & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
-    & \quad \quad \forall \npair[n']{c} \in \stdkr(W) \ldotp \\
     & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
-    & \quad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
+    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
   \end{aligned}

### 2016-11-15 capability-machine.tex
Some minor changes I forgot to commit.
Outcommented PreviewMacro
Oops, fix compilation.
added semantics for restrict
Added a couple of small macros
typos

@@
 \usepackage{multirow}
+
+%% make latex preview-mode work with natbib...
+\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
+%\PreviewMacro*{\cofes{}} This brings compiling for non-auctex to a grinding halt.
 
@@
 \newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
+\newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
 \newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
@@
        &   & \lea{\lv}{\rv} \mid\move{\lv}{\rv} \mid \load{\lv}{\hv} \mid \store{\hv}{\rv} \mid  \\
-       &   & \restrict{\lv}{\rv}{\rv} \mid \subseg{\lv}{\rv}{\rv} \mid \plus{\lv}{\rv}{\rv}
+       &   & \restricttwo{\lv}{\rv} \mid \subseg{\lv}{\rv}{\rv} \mid \plus{\lv}{\rv}{\rv}
 \end{array}$$
@@
                            \arraycolsep=0pt
-                           \begin{array}{l}
+                           \begin{array}[t]{l}
                              \text{if $\memreg(\pcreg) = \stdcap$}\\
@@
                                                \arraycolsep=0pt
-                                               \begin{array}{l}
+                                               \begin{array}[t]{l}
                                                  \text{if $\nonZero{\memreg(\rv)}$} 
@@
                                                \arraycolsep=0pt
-                                               \begin{array}{l}
+                                               \begin{array}[t]{l}
                                                  \text{if }\memreg(r_2) = \stdcap = \var{c} \\
@@
                                                \arraycolsep=0pt
-                                               \begin{array}{l}
+                                               \begin{array}[t]{l}
                                                  \text{if }\memreg(r_1) = \stdcap = \var{c} \\
@@
                                                  \arraycolsep=0pt
-                                                 \begin{array}{l}
+                                                 \begin{array}[t]{l}
                                                    \text{if either $n = \rv$ or $\rv = \refreg{r_2}$ and $n = \memreg(r_2)$} \\
@@
  % In the M-Machine, lea checks whether the pointer stay within the allowed range in the pointer.
-  \sem{\restrict{\refreg{r_1}}{\rv_1}{\rv_2}}           & =
+  \sem{\restricttwo{\refreg{r}}{\rv}}           & =
                                              \begin{cases}
-                                               \stdUpdatePc{\updateReg{r_1}{\var{c}}}  &
+                                               \stdUpdatePc{\updateReg{r}{\var{c}}}  &
                                                  \arraycolsep=0pt
-                                                 \begin{array}{l}
-                                                   \text{if $\memreg(\rv_1) = c'$}\\
-                                                   \quad\text{and $c' = \stdcap[\permp]$}\\
-                                                   \quad\text{and either $\rv_2 = n$ or $\memreg(\rv_2) = n$}\\
+                                                 \begin{array}[t]{l}
+                                                   \text{if $\memreg(r) = \stdcap[\permp]$}\\
+                                                   \quad\text{and either $\rv = n$ or $\memreg(\rv) = n$}\\
                                                    \quad\text{and in either case $n \in \ints$}\\
@@
                                                    \quad\text{and $\var{newPermPair} \sqsubseteq \permp$}\\
-                                                   \quad\text{and $c = (\var{newPerm},\start,\addrend,\addr)$}
+                                                   \quad\text{and $c = (\var{newPermPair},\start,\addrend,\addr)$}
                                                  \end{array}\\
@@
                                                             \arraycolsep=0pt
-                                                            \begin{array}{l}
+                                                            \begin{array}[t]{l}
                                                               \text{if for $i \in \{1,2\}$}\\
@@
                                                           \end{cases}\\
+  \sem{\subseg{\refreg{r}}{\rv_1}{\rv_2}} & = 
+                                            \begin{cases}
+                                              \stdUpdatePc{\updateReg{r}{\var{c}}} &
+                                              \arraycolsep=0pt
+                                              \begin{array}[t]{l}
+                                                \text{if $\memreg{r} = \stdcap[\permp]$} \\
+                                                \quad\text{and for $i \in \{1,2\}$}\\
+                                                \quad\text{$n_i = \rv_i$ or $n_i = \memreg(\rv_i)$}\\
+                                                \quad\text{and in either case $n_i \in \ints$}\\
+                                                \quad\text{and $\start \leq n_1 \leq n_2 \leq \addrend$}\\
+                                                \quad\text{and $c = (\permp,n_2,n_2,\addr)$}
+                                              \end{array} \\
+                                              \failed & \text{otherwise}
+                                            \end{cases}
+ \\
   \sem{\isptr{\lv}{\rv}} & = \undefined \\ 
-  \sem{\setptr{\lv}{\rv}} & = \undefined \\ 
-  \sem{\subseg{\lv}{\rv}{\rv}} & = \undefined 
+  \sem{\setptr{\lv}{\rv}} & = \undefined 
 \end{align*}
@@
 Whether a $\restrict{}{}{}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above.
+
+Define the following simple macros:
+
+\begin{align*}
+  \restrict{r_1}{r_2}{r_3} \; r_4 &\defeq
+                                    \begin{aligned}[t]
+                                      & \move{r_1}{r_2} \\
+                                      & \restrict{r_1}{r_3}{r_4}
+                                    \end{aligned} \\
+  \subseg{r_1}{r_2}{r_3} \; r_4   &\defeq
+                                    \begin{aligned}[t]
+                                      & \move{r_1}{r_2} \\
+                                      & \subseg{r_1}{r_3}{r_4}
+                                    \end{aligned}\\
+  \lea{r_1}{r_2} \; r_3           &\defeq
+                                    \begin{aligned}[t]
+                                      & \move{r_1}{r_2} \\
+                                      & \lea{r_1}{r_3}
+                                    \end{aligned}\\
+  \store{r}{n} &\defeq
+                   \begin{aligned}[t]
+                     & \move{r_i}{n} \qquad \text{ where $r_i$ unused register}\\
+                     & \store{r}{r_i}
+                   \end{aligned}
+\end{align*}
 
@@
               \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
-            { W' \futurestr W }
+            { W' \futurewk W }
 \and
@@
 \end{align*}
-\lau{In the above $H$ definitions, $\hat{W}$ is not a world, so there is a type problem that needs to be fixed. Maybe solve the recursive domain equations wrt the worlds instead.}
 
@@
 \end{align*}
-$\localReg$ are the regions that local capabilities may govern - that is permanent and global regions.
+$\localReg$ are the regions that local capabilities may govern - that is permanent and temporary regions.
 $\globalReg$ are the regions that global capabilities may govern - that is permanent regions.
@@
 \end{align*}
-\lau{I am in doubt whether it should be $\iota^{\var{pwl}}$ in the read condition.}
 
@@
 We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
-\dominique{this sentence doesn't parse :)}
 \begin{definition}[Preordered \cofe{}\ construction: Function]
@@
 \end{verbatim}
-\lau{I have inserted some line breaks for readability. I am not sure what is going on here (the parenthesis after y is unmatched.)}
+\lau{I have inserted some line breaks for readability. not sure what is going on here (the parenthesis after y is unmatched.)}
 \section{Related reading}

### 2016-11-16 capability-machine.tex
Added heap satisfaction
Added an akward like example

@@
 \usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
-%\PreviewMacro*{\cofes{}} This brings compiling for non-auctex to a grinding halt.
 
@@
 
+\subsection{Variant of the ``awkward'' example}
+Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
+\begin{verbatim}
+let h = fun adv =>
+          let x = 1 in
+          let g = fun adv =>
+                    x := 0;
+                    adv(g);
+                    x := 1;
+                    adv(g);
+                    assert(x == 1) in
+          g(g)
+\end{verbatim}
+In Dreyer's example, the issue arises as the context can provide the result of evaluating $\mathtt{e_1}$ (the equivalent here is \texttt{g})) to \texttt{f} (the equivalent here is \texttt{adv}).
+In our system, the context cannot capture this, so in the above program, \texttt{adv} is given explicit access to \texttt{g}.
+The lemma we would want to prove here is that the assertion does not fail.
+
 \subsection{Ticket Dispenser}
@@
 \begin{align*}
-  \erase{W}{S} \defeq \lambda r \ldotp 
+  \lfloor W \rfloor_S \defeq \lambda r \ldotp 
   \begin{cases}
@@
   \end{cases}
+\end{align*}
+
+Heap satisfaction:
+\begin{align*}
+  &\heapSat[\hs]{n}{W} \\
+  &\qquad\text{ iff }\\
+  &\begin{aligned}[t]
+    &\exists R : \dom(\erase{W}{\perma,\temp}) \rightarrow \HeapSegments \ldotp \\
+    &\quad \hs = \uplus_{r\in\dom(\erase{W}{\perma,\temp})}R(r) \land \\
+    &\quad \forall r \in \dom(\erase{W}{\perma,\temp}) \ldotp\\
+    &\qquad \forall n'<n \ldotp \\
+    &\qquad \quad \exists H,s \ldotp\\
+    &\qquad \qquad W(r) = (\_,s,\_,H) \land \\
+    &\qquad \qquad \npair[n']{R(r)} \in H(s)(\xi^{-1}(W))\\
+  \end{aligned}
 \end{align*}
@@
 \end{align*}
-\end{definition}
+o\end{definition}
 

### 2016-11-16 capability-machine.tex
Added new projection instructions for capability fields

@@
 \newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
-\newcommand{\setptr}[2]{\twoinstr{setptr}{#1}{#2}}
+\newcommand{\geta}[2]{\twoinstr{geta}{#1}{#2}}
+\newcommand{\getb}[2]{\twoinstr{getb}{#1}{#2}}
+\newcommand{\gete}[2]{\twoinstr{gete}{#1}{#2}}
+\newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
 \newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
@@
 \rv    &::=& n \mid \lv \\
-i      &::=& \fail \mid \halt \mid 
-             \jmp{\lv} \mid \jnz{\lv}{\rv} \mid
-             \isptr{\lv}{\rv} \mid \setptr{\lv}{\rv} \mid \\
-       &   & \lea{\lv}{\rv} \mid\move{\lv}{\rv} \mid \load{\lv}{\hv} \mid \store{\hv}{\rv} \mid  \\
-       &   & \restricttwo{\lv}{\rv} \mid \subseg{\lv}{\rv}{\rv} \mid \plus{\lv}{\rv}{\rv}
+i      &::=& 
+             \jmp{\lv} \mid 
+             \jnz{\lv}{\rv} \mid
+             \move{\lv}{\rv} \mid 
+             \load{\lv}{\hv} \mid 
+             \store{\hv}{\rv} \mid  \\
+       &   & \plus{\lv}{\rv}{\rv} \mid 
+             \lea{\lv}{\rv} \mid 
+             \restricttwo{\lv}{\rv} \mid 
+             \subseg{\lv}{\rv}{\rv} \mid  \\
+       &   & \isptr{\lv}{\rv} \mid 
+             \getp{\lv}{\lv} \mid 
+             \getb{\lv}{\lv} \mid
+             \gete{\lv}{\lv} \mid
+             \geta{\lv}{\lv} \mid \\
+       &   & \fail \mid
+             \halt \mid 
 \end{array}$$
@@
  \\
-  \sem{\isptr{\lv}{\rv}} & = \undefined \\ 
-  \sem{\setptr{\lv}{\rv}} & = \undefined 
+  \sem{\geta{\refreg{r_1}}{\refreg{r_2}}} & = 
+                          \begin{cases}
+                            \stdUpdatePc{\updateReg{r_1}{\addr}} &
+                            \arraycolsep=0pt
+                            \begin{array}[t]{l}
+                              \text{if $\memreg{r_2} = \stdcap$}
+                            \end{array} \\
+                            \failed & \text{otherwise}
+                          \end{cases}
+  \\
+  \sem{\getb{\refreg{r_1}}{\refreg{r_2}}} & = 
+                          \begin{cases}
+                            \stdUpdatePc{\updateReg{r_1}{\start}} &
+                            \arraycolsep=0pt
+                            \begin{array}[t]{l}
+                              \text{if $\memreg{r_2} = \stdcap$}
+                            \end{array} \\
+                            \failed & \text{otherwise}
+                          \end{cases}
+  \\
+  \sem{\gete{\refreg{r_1}}{\refreg{r_2}}} & = 
+                          \begin{cases}
+                            \stdUpdatePc{\updateReg{r_1}{\addrend}} &
+                            \arraycolsep=0pt
+                            \begin{array}[t]{l}
+                              \text{if $\memreg{r_2} = \stdcap$}
+                            \end{array} \\
+                            \failed & \text{otherwise}
+                          \end{cases}
+  \\
+  \sem{\getp{\refreg{r_1}}{\refreg{r_2}}} & = 
+                          \begin{cases}
+                            \stdUpdatePc{\updateReg{r_1}{\perm}} & \text{if $\memreg{r_2} = \stdcap$} \\
+                            \failed & \text{otherwise}
+                          \end{cases}
+  \\
+%getl?
+  \sem{\isptr{\refreg{r}}{\rv}} & =  
+                          \begin{cases}
+                            \stdUpdatePc{\updateReg{r_1}{1}} & \text{if $\memreg{\rv} = \stdcap$} \\
+                            \stdUpdatePc{\updateReg{r_1}{0}} & \text{otherwise} 
+                          \end{cases}
 \end{align*}
@@
 \label{subsec:example-loc-cap}
-In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate with the return value 1.
+In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1.
+The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}).
 Notice that an assertion will never terminate in a $\failed$ state.
@@
           adv();
-          return 0
+          return
 \end{verbatim}
@@
           adv(0);
-          return 0
+          return
 \end{verbatim}
@@
 The lemma we would want to prove here is that the assertion does not fail.
+
+To show this, we will probably need public/private transitions.
 

### 2016-11-17 capability-machine.tex
Added getp and getl

@@
 \newcommand{\encode}{\plainfun{encode}{}}
-\newcommand{\encodePerm}{\plainfun{encodePerm}{}}
-\newcommand{\decodePerm}{\plainfun{decodePerm}{}}
+\newcommand{\encodePerm}{\mathit{encodePerm}}
+\newcommand{\encodePermPair}{\plainfun{encodePermPair}{}}
+\newcommand{\encodeLoc}{\mathit{encodeLoc}{}}
+\newcommand{\decodePermPair}{\plainfun{decodePerm}}
 \newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}
@@
 \newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
+\newcommand{\getl}[2]{\twoinstr{getl}{#1}{#2}}
 \newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
@@
 Notation:
-$$\begin{array}{rcl}
+\[
+\begin{array}{rcl}
 i       &\in& \Instrs \\
@@
 n       &\in& \ints\\
-\end{array}$$
+\end{array}
+\]
 Further definitions:
-$$\begin{array}{rcl}
+\[
+\begin{array}{rcl}
 \lv    &::=& \refreg{r} \\
@@
              \getp{\lv}{\lv} \mid 
+             \getl{\lv}{\lv} \mid 
              \getb{\lv}{\lv} \mid
@@
              \halt \mid 
-\end{array}$$
+\end{array}
+\]
 
@@
 \end{align*}
-Assume an $\encodePerm$ function that encodes a permission as an integer:
-\begin{align*}
-\encodePerm &: (\Perms \times \Globals) \fun \ints
-\end{align*}
-Further, assume an inverse function, $\decodePerm$, that decodes permissions
-\[
-  \decodePerm : \ints \fun (\Perms \times \Globals)
+Assume an $\encodePerm$, $\encodeLoc$, and $\encodePermPair$ function that encodes a permissions, locality, and permission pair, respectively, as an integer:
+\begin{align*}
+\encodePerm &: \Perms \fun \ints \\
+\encodeLoc &: \Globals \fun \ints \\
+\encodePermPair &: (\Perms \times \Globals) \fun \ints \\
+\end{align*}
+Further, assume an inverse function, $\decodePermPair$, that decodes permissions
+\[
+  \decodePermPair : \ints \fun (\Perms \times \Globals)
 \]
@@
                                                    \quad\text{and in either case $n \in \ints$}\\
-                                                   \quad\text{and $\var{newPermPair} = \decodePerm(n)$}\\
+                                                   \quad\text{and $\var{newPermPair} = \decodePermPair(n)$}\\
                                                    \quad\text{and $\var{newPermPair} \sqsubseteq \permp$}\\
@@
                           \begin{cases}
-                            \stdUpdatePc{\updateReg{r_1}{\perm}} & \text{if $\memreg{r_2} = \stdcap$} \\
+                            \stdUpdatePc{\updateReg{r_1}{\encodePerm(\perm)}} &
+                            \arraycolsep=0pt
+                            \begin{array}[t]{l}
+                              \text{if $\memreg{r_2} = \stdcap$}
+                            \end{array} \\
                             \failed & \text{otherwise}
@@
   \\
-%getl?
+  \sem{\getl{\refreg{r_1}}{\refreg{r_2}}} & = 
+                          \begin{cases}
+                            \stdUpdatePc{\updateReg{r_1}{\encodeLoc(\gl)}} &
+                            \arraycolsep=0pt
+                            \begin{array}[t]{l}
+                              \text{if $\memreg{r_2} = \stdcap$}
+                            \end{array} \\
+                            \failed & \text{otherwise}
+                          \end{cases}
+  \\
   \sem{\isptr{\refreg{r}}{\rv}} & =  
@@
 \subsection{Another simple well-bracketedness/local capability example}
-Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability.
+Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability (but \texttt{let x = ...} produces a global capability). 
 \begin{verbatim}
 let f = fun adv =>
+          let x = 1 in
           let l = 0 in
-          let x = 1 in
           adv(l);
@@
           let x = 1 in
-          let g = fun adv =>
+          let g = fun g' =>
+                    x := 1;
+                    adv(g');
                     x := 0;
-                    adv(g);
-                    x := 1;
-                    adv(g);
-                    assert(x == 1) in
-          g(g)
+                    adv();
+                    assert(x == 0) in
+          g(g);
+          return
 \end{verbatim}
@@
   \observations (W) \defeq & \lambda W \ldotp 
+                               \{ \npair{(\reg,\hs)} \mid
                              \begin{aligned}[t]
-                               \{ \npair{(\reg,\hs)} \mid & \forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
-                                                          & \quad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'} \}
+                               & \forall \heap_f, \heap', i \leq n \ldotp \\
+                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                               & \qquad \Rightarrow
+                               \begin{aligned}[t]
+                                 & \exists W' \futurestr W \ldotp\exists \hs' \ldotp\\
+                                 & \quad \heap' = \hs' \uplus \heap_f \land \\ 
+                                 & \quad \heapSat[\hs']{n-i}{W'} \}
+                               \end{aligned}
                              \end{aligned}

### 2016-11-17 capability-machine.tex
added revoked heap segment to observation relation
revised awkward example and correctness lemma

@@
           adv(0);
-          assert(l == 1)
+          assert(!l == 1)
 \end{verbatim}
@@
           adv();
-          assert(x == 1);
+          assert(!x == 1);
           x := 2;
@@
           adv(l);
-          assert(x == 1);
+          assert(!x == 1);
           x := 2;
@@
 \begin{verbatim}
-let h = fun adv =>
-          let x = 1 in
-          let g = fun g' =>
-                    x := 1;
-                    adv(g');
-                    x := 0;
-                    adv();
-                    assert(x == 0) in
-          g(g);
-          return
+g = fun _ => let x = 0 in
+               fun f =>
+                 x:=0;
+                 f();
+                 x:=1;
+                 f();
+                 assert(x==1)
 \end{verbatim}
-In Dreyer's example, the issue arises as the context can provide the result of evaluating $\mathtt{e_1}$ (the equivalent here is \texttt{g})) to \texttt{f} (the equivalent here is \texttt{adv}).
-In our system, the context cannot capture this, so in the above program, \texttt{adv} is given explicit access to \texttt{g}.
-The lemma we would want to prove here is that the assertion does not fail.
-
-To show this, we will probably need public/private transitions.
+\begin{lemma}
+  \newcommand{\adv}{\var{adv}}
+  \newcommand{\flag}{\var{flag}}
+  For all $n \in \nats$ 
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
+    c_{g} & \defeq ((\entry,\glob),\start_g,\addrend_g,\start_g)
+  \end{align*}
+if
+  \[
+    (\reg_0\update{\pcreg}{c_{\adv}}
+          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus h_f) \step[n] (\halted,h')
+  \]
+where 
+\begin{align*}
+  &\dom(hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+  &\dom(hs_g) = [\start_g,\addrend_g] \\
+  &\dom(hs_\flag) = [\flag,\flag] \\
+  &\heapSat[\hs_{\adv}]{n}{\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}}
+\end{align*}
+and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$, 
+then
+\[
+  h'(\flag) = 0
+\]
+\end{lemma}
 
@@
                                \begin{aligned}[t]
-                                 & \exists W' \futurestr W \ldotp\exists \hs' \ldotp\\
-                                 & \quad \heap' = \hs' \uplus \heap_f \land \\ 
+                                 & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
+                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \heap_f \land \\ 
                                  & \quad \heapSat[\hs']{n-i}{W'} \}

### 2016-11-18 capability-machine.tex
Added lemmas we expect to hold
changed the awkward example slightly

@@
 \newcommand{\future}{\mathbin{\sqsupseteq}}
-\newcommand{\futurewk}{\mathbin{\sqsupseteq}_{\var{wk}}}
-\newcommand{\futurestr}{\mathbin{\sqsupseteq}_{\var{str}}}
+\newcommand{\futurewk}{\mathbin{\sqsupseteq}^{\var{pub}}}
+\newcommand{\futurestr}{\mathbin{\sqsupseteq}^{\var{priv}}}
 \newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
@@
 \end{array}
+\]
+Further define $\reg_0 \in \Regs$ such that
+\[
+  \forall r \in \RegName \ldotp \reg_0(r) = 0
 \]
@@
 \end{verbatim}
-\begin{lemma}
+\begin{lemma}\forcenewline
   \newcommand{\adv}{\var{adv}}
@@
     (\reg_0\update{\pcreg}{c_{\adv}}
-          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus h_f) \step[n] (\halted,h')
+          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus h_f) \step[n] (\halted,h')
   \]
@@
   &\dom(hs_\flag) = [\flag,\flag] \\
-  &\heapSat[\hs_{\adv}]{n}{\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}}
-\end{align*}
-and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$, 
+  &\heapSat[\hs_{\adv} \uplus \hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}, 1 \mapsto\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}]}
+\end{align*}
+and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$ as well as an entry capability for $\malloc$, 
 then
@@
 
+\section{Logical Relation'}
+\subsection{Worlds}
+Assume $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
+\begin{align*}
+  \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
+                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
+                                                                & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}}) \\ \\
+  \xi_{\var{str}} :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
+                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
+                                                                & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}})
+\end{align*}
+We now define the regions to be
+\begin{align*}
+  \Regions = & \{\revoked\} \uplus \\
+             & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \uplus \\
+             & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))
+\end{align*}
+And the worlds are
+\[
+  \Worlds = \RegionNames \finparfun \Regions
+\]
+Where $\RegionNames = \nats$.
+
+Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system:
+\begin{mathpar}
+  \inferrule{  (s_1,s_2) \in \phi_2 \\
+               (\phi_1,H_1) = (\phi_2,H_2)}
+            {  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) }
+\end{mathpar}
+our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
+\begin{mathpar}
+  \inferrule{ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
+               v_1 = v_2 }
+            { (v_2,s_2,\phi_2,H_2) \futurewk (v_1,s_1,\phi_1,H_1) }
+\and
+  \inferrule{ }
+            { \revoked \futurewk \revoked }
+\end{mathpar}
+For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
+\begin{mathpar}
+  \inferrule{ r_2 \futurewk  r_1 }
+            { r_2 \futurestr r_1 }
+\and
+  \inferrule{ }
+            { \revoked \futurestr (\temp,s,\phi,H)}
+\end{mathpar}
+
+We now get two future world relations $\futurewk$ and $\futurestr$ are defined as one would expect.
+That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
+\begin{mathpar}
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+              \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
+            { W' \futurewk W }
+\and
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+              \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
+            { W' \futurestr W }
+\end{mathpar}
+
+Erase all but a set of views:
+\begin{align*}
+  \lfloor W \rfloor_S \defeq \lambda r \ldotp 
+  \begin{cases}
+    W(r) & W(r).v \in S\\
+    \bot & \text{otherwise}
+  \end{cases}
+\end{align*}
+
+Heap satisfaction:
+\begin{align*}
+  &\heapSat[\hs]{n}{W} \\
+  &\qquad\text{ iff }\\
+  &\begin{aligned}[t]
+    &\exists R : \dom(\erase{W}{\perma,\temp}) \rightarrow \HeapSegments \ldotp \\
+    &\quad \hs = \uplus_{r\in\dom(\erase{W}{\perma,\temp})}R(r) \land \\
+    &\quad \forall r \in \dom(\erase{W}{\perma,\temp}) \ldotp\\
+    &\qquad \forall n'<n \ldotp \\
+    &\qquad \quad \exists H,s \ldotp\\
+    &\qquad \qquad W(r) = (\_,s,\_,H) \land \\
+    &\qquad \qquad \npair[n']{R(r)} \in H(s)(\xi^{-1}(W))\\
+  \end{aligned}
+\end{align*}
+
+Standard regions for when writing locally is permitted:
+\begin{align*}
+  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
+  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
+  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
+    \end{aligned}
+  \right\}
+\end{align*}
+Revoking all temporary regions:
+\begin{align*}
+  \revokeTemp{} & : \Worlds \fun \Worlds \\
+  \revokeTemp{W} & \defeq \lambda r \ldotp 
+                   \begin{cases}
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
+                     W(r)                & \text{otherwise}
+                   \end{cases}
+\end{align*}
+\newcommand{\wrev}[1]{#1_{\var{rev}}}
+Let
+\[
+\revokeTemp(W) = \wrev{W}
+\]
+
+Standard regions for when write local is not allowed:
+\begin{align*}
+  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
+  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
+    \begin{cases}
+      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
+      \bot & \text{otherwise}
+\end{cases}
+\\ \\
+  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \\
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
+    \end{aligned}
+  \right\}
+\end{align*}
+
+For convenience define
+\begin{align*}
+  \localReg(W) & \defeq \dom(\erase{W}{\perm,\temp}) \\
+  \globalReg(W) & \defeq \dom(\erase{W}{\perm})
+\end{align*}
+$\localReg$ are the regions that local capabilities may govern - that is permanent and temporary regions.
+$\globalReg$ are the regions that global capabilities may govern - that is permanent regions.
+Now define the following function
+\[
+  \var{localityReg}(\gl,W) \defeq 
+  \begin{cases}
+    \localReg(W) & \text{if } \gl = \local \\
+    \globalReg(W) & \text{if } \gl = \glob
+  \end{cases}
+\]
+
+\begin{align*}
+    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+    & \writeCond{}(\iota,\gl)(W) =  \\
+    & \quad \begin{aligned}[t]
+              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+              & \;\qquad W(r)\nsupeq[n-1] \iota_{\start',\addrend'}(W(r).v) \}
+            \end{aligned}
+\end{align*}
+
+\begin{align*}
+    & \readCond{}(\gl)(W) =  \\
+    & \quad \begin{aligned}[t]
+              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+                        & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+                        & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
+            \end{aligned}
+\end{align*}
+
+\begin{align*}
+  & \execCond{}(\gl)(W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad\forall a \in [\start,\addrend] \ldotp\\
+      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+\begin{align*}
+  & \entryCond{}(\gl)(W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+Now define the value relation as follows:
+\begin{align*}
+  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
+  &\stdvr\defeq \lambda \; W \ldotp 
+           \begin{aligned}[t]
+             & \{ \npair{i} \mid i \in \ints \} 
+             \union \\
+             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
+             \union \\
+             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
+             \union \\
+             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \}
+             \union \\
+             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \}
+             \union \\
+             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
+             \union \\
+             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
+             & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
+             \union \\
+             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \land\\
+             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
+             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
+             \union \\
+             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
+             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \land\\
+             & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
+             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
+             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
+           \end{aligned}
+\end{align*}
+
+\begin{align*}
+  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
+  \observations (W) \defeq & \lambda W \ldotp 
+                               \{ \npair{(\reg,\hs)} \mid
+                             \begin{aligned}[t]
+                               & \forall \heap_f, \heap', i \leq n \ldotp \\
+                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                               & \qquad \Rightarrow
+                               \begin{aligned}[t]
+                                 & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
+                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \heap_f \land \\ 
+                                 & \quad \heapSat[\hs']{n-i}{W'} \}
+                               \end{aligned}
+                             \end{aligned}
+\end{align*}
+
+\begin{align*}
+  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
+  \stdrr \defeq & \lambda W \ldotp
+                     \begin{aligned}[t]
+                       \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                                            & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
+                     \end{aligned}
+\end{align*}
+
+\begin{align*}
+  \stder : & \Worlds \nefun \UPred{\Words} \\
+  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
+  \begin{aligned}[t]
+    & \forall n' \leq n \ldotp\\
+    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
+    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
+    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
+  \end{aligned}
+\end{align*}
+
+\subsection{Lemmas}
+We expect the following lemmas to hold true:
+\subsubsection{Heap Satisfaction}
+\begin{lemma}[Downwards closure]
+  \begin{align*}
+    &\forall \hs, n' leq n, W \ldotp \\
+    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Public monotonicity]
+  \begin{align*}
+    & \forall \hs, n, W' \futurewk W \ldotp \\
+    & \quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n}{W'}
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Private monotonicity-like property]
+  \begin{align*}
+    & \forall \hs, n, W, W' \ldotp \\
+    & \quad  \heapSat[\hs]{n}{W} \land W' \futurewk \wrev{W} \futurestr W \\
+    & \qquad \Rightarrow \exists \hs', \hs_f \ldotp \hs = \hs' \uplus \hs_f \land \heapSat[\hs']{n}{W'}
+  \end{align*}
+\end{lemma}
+
+\subsubsection{Future world}
+\begin{lemma}
+  For all $W$, $W'$, and $W''$
+  \begin{itemize}
+  \item $W' \futurewk W \Rightarrow W' \futurestr W$
+  \item $\wrev{W} \futurestr W$
+  \item $W \futurestr \wrev{W} \Rightarrow W \futurewk \wrev{W}$
+  \end{itemize}
+  Combined with transitivity of $\futurestr$, the above gives us:
+  \begin{itemize}
+  \item $W'' \futurestr W' \land W' \futurewk W \Rightarrow W'' \futurestr W$
+  \item $W'' \futurewk W' \land W' \futurestr W \Rightarrow W'' \futurestr W$
+  \end{itemize}
+\end{lemma}
+
+\subsubsection{Value relation}
+\begin{lemma}[Global capabilities monotone wrt $\futurestr$]
+  \begin{align*}
+    & \forall n, \perm, \start, \addrend, \addr, W \ldotp \\
+    & \quad  \npair{\stdcap[(\perm,\glob)]} \in \stdvr(W) \\
+    & \qquad \Rightarrow \npair{\stdcap[(\perm,\glob)]} \in \stdvr(\wrev{W}) 
+  \end{align*}
+\end{lemma}
+
+\section{Old TD example}
 \subsection{Ticket Dispenser}
@@
 
-\section{Logical Relation'}
-\subsection{Worlds}
-Assume $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
-\begin{align*}
-  \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
-                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                                                & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}}) \\ \\
-  \xi_{\var{str}} :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
-                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                                                & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}})
-\end{align*}
-We now define the regions to be
-\begin{align*}
-  \Regions = & \{\revoked\} \uplus \\
-             & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \uplus \\
-             & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))
-\end{align*}
-And the worlds are
-\[
-  \Worlds = \RegionNames \finparfun \Regions
-\]
-Where $\RegionNames = \nats$.
-
-Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system:
-\begin{mathpar}
-  \inferrule{  (s_1,s_2) \in \phi_2 \\
-               (\phi_1,H_1) = (\phi_2,H_2)}
-            {  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) }
-\end{mathpar}
-our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
-\begin{mathpar}
-  \inferrule{ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
-               v_1 = v_2 }
-            { (v_2,s_2,\phi_2,H_2) \futurewk (v_1,s_1,\phi_1,H_1) }
-\and
-  \inferrule{ }
-            { \revoked \futurewk \revoked }
-\end{mathpar}
-For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
-\begin{mathpar}
-  \inferrule{ r_2 \futurewk  r_1 }
-            { r_2 \futurestr r_1 }
-\and
-  \inferrule{ }
-            { \revoked \futurestr (\temp,s,\phi,H)}
-\end{mathpar}
-
-We now get two future world relations $\futurewk$ and $\futurestr$ are defined as one would expect.
-That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
-\begin{mathpar}
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-              \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
-            { W' \futurewk W }
-\and
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-              \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
-            { W' \futurestr W }
-\end{mathpar}
-
-Erase all but a set of views:
-\begin{align*}
-  \lfloor W \rfloor_S \defeq \lambda r \ldotp 
-  \begin{cases}
-    W(r) & W(r).v \in S\\
-    \bot & \text{otherwise}
-  \end{cases}
-\end{align*}
-
-Heap satisfaction:
-\begin{align*}
-  &\heapSat[\hs]{n}{W} \\
-  &\qquad\text{ iff }\\
-  &\begin{aligned}[t]
-    &\exists R : \dom(\erase{W}{\perma,\temp}) \rightarrow \HeapSegments \ldotp \\
-    &\quad \hs = \uplus_{r\in\dom(\erase{W}{\perma,\temp})}R(r) \land \\
-    &\quad \forall r \in \dom(\erase{W}{\perma,\temp}) \ldotp\\
-    &\qquad \forall n'<n \ldotp \\
-    &\qquad \quad \exists H,s \ldotp\\
-    &\qquad \qquad W(r) = (\_,s,\_,H) \land \\
-    &\qquad \qquad \npair[n']{R(r)} \in H(s)(\xi^{-1}(W))\\
-  \end{aligned}
-\end{align*}
-
-Standard regions for when writing locally is permitted:
-\begin{align*}
-  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
-  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
-    \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
-    \end{aligned}
-  \right\}
-\end{align*}
-Revoking all temporary regions:
-\begin{align*}
-  \revokeTemp{} & : \Worlds \fun \Worlds \\
-  \revokeTemp{W} & \defeq \lambda r \ldotp 
-                   \begin{cases}
-                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
-                     W(r)                & \text{otherwise}
-                   \end{cases}
-\end{align*}
-Standard regions for when write local is not allowed:
-\begin{align*}
-  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
-  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
-    \begin{cases}
-      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
-      \bot & \text{otherwise}
-\end{cases}
-\\ \\
-  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
-    \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
-    \end{aligned}
-  \right\}
-\end{align*}
-
-For convenience define
-\begin{align*}
-  \localReg(W) & \defeq \dom(\erase{W}{\perm,\temp}) \\
-  \globalReg(W) & \defeq \dom(\erase{W}{\perm})
-\end{align*}
-$\localReg$ are the regions that local capabilities may govern - that is permanent and temporary regions.
-$\globalReg$ are the regions that global capabilities may govern - that is permanent regions.
-Now define the following function
-\[
-  \var{localityReg}(\gl,W) \defeq 
-  \begin{cases}
-    \localReg(W) & \text{if } \gl = \local \\
-    \globalReg(W) & \text{if } \gl = \glob
-  \end{cases}
-\]
-
-\begin{align*}
-    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
-    & \writeCond{}(\iota,\gl)(W) =  \\
-    & \quad \begin{aligned}[t]
-              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-              & \;\qquad W(r)\nsupeq[n-1] \iota_{\start',\addrend'}(W(r).v) \}
-            \end{aligned}
-\end{align*}
-
-\begin{align*}
-    & \readCond{}(\gl)(W) =  \\
-    & \quad \begin{aligned}[t]
-              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-                        & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-                        & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
-            \end{aligned}
-\end{align*}
-
-\begin{align*}
-  & \execCond{}(\gl)(W) = \\
-  & \quad
-    \begin{aligned}[t]
-      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
-      & \quad \forall W' \future W \ldotp\\
-      & \qquad\forall a \in [\start,\addrend] \ldotp\\
-      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
-    \end{aligned} \\
-  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
-\end{align*}
-
-\begin{align*}
-  & \entryCond{}(\gl)(W) = \\
-  & \quad
-    \begin{aligned}[t]
-      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
-      & \quad \forall W' \future W \ldotp\\
-      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
-    \end{aligned} \\
-  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
-\end{align*}
-
-Now define the value relation as follows:
-\begin{align*}
-  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
-  &\stdvr\defeq \lambda \; W \ldotp 
-           \begin{aligned}[t]
-             & \{ \npair{i} \mid i \in \ints \} 
-             \union \\
-             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
-             \union \\
-             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
-             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
-             \union \\
-             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
-             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \}
-             \union \\
-             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
-             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \}
-             \union \\
-             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
-             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
-             \union \\
-             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
-             & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
-             \union \\
-             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
-             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \land\\
-             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
-             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
-             \union \\
-             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
-             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \land\\
-             & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
-             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
-             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
-           \end{aligned}
-\end{align*}
-
-\begin{align*}
-  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-  \observations (W) \defeq & \lambda W \ldotp 
-                               \{ \npair{(\reg,\hs)} \mid
-                             \begin{aligned}[t]
-                               & \forall \heap_f, \heap', i \leq n \ldotp \\
-                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
-                               & \qquad \Rightarrow
-                               \begin{aligned}[t]
-                                 & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
-                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \heap_f \land \\ 
-                                 & \quad \heapSat[\hs']{n-i}{W'} \}
-                               \end{aligned}
-                             \end{aligned}
-\end{align*}
-
-\begin{align*}
-  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
-  \stdrr \defeq & \lambda W \ldotp
-                     \begin{aligned}[t]
-                       \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
-                                            & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
-                     \end{aligned}
-\end{align*}
-
-\begin{align*}
-  \stder : & \Worlds \nefun \UPred{\Words} \\
-  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
-  \begin{aligned}[t]
-    & \forall n' \leq n \ldotp\\
-    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
-    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
-    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
-  \end{aligned}
-\end{align*}
-
 \section{Logical Relation}

### 2016-11-18 capability-machine.tex
Lau and Dominique meeting

@@
 \begin{align*}
-  &\heapSat[\hs]{n}{W} \\
-  &\qquad\text{ iff }\\
-  &\begin{aligned}[t]
+  &\heapSat[\hs]{n}{W} 
+  \text{ iff }
+  \left\{\begin{aligned}
     &\exists R : \dom(\erase{W}{\perma,\temp}) \rightarrow \HeapSegments \ldotp \\
@@
     &\qquad \qquad \npair[n']{R(r)} \in H(s)(\xi^{-1}(W))\\
-  \end{aligned}
-\end{align*}
+  \end{aligned}\right.
+\end{align*}
+\dominique{quantification over $n' < n$ needed?}
 
@@
 \end{align*}
-\newcommand{\wrev}[1]{#1_{\var{rev}}}
-Let
-\[
-\revokeTemp(W) = \wrev{W}
-\]
+\newcommand{\wrev}[1]{\revokeTemp{#1}}
 
@@
   \begin{align*}
-    &\forall \hs, n' leq n, W \ldotp \\
+    &\forall \hs, n' \leq n, W \ldotp \\
     &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
@@
   \item $\wrev{W} \futurestr W$
-  \item $W \futurestr \wrev{W} \Rightarrow W \futurewk \wrev{W}$
   \end{itemize}
@@
 \begin{specification}[Malloc v.2]
+  $c_\malloc$ satisfies the specification for malloc iff
   \begin{align*}
@@
     &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
-    &\qquad \exists i, \iota_{\malloc} \ldotp \\
-    &\qquad \quad W = [i \mapsto \iota_{\malloc}] \land \\
-    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\ % This
+    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
+    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
+%    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\ % This
     &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
@@
     &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
-    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\ %and this needs to change with ABI.
+    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
     &\qquad \quad \Rightarrow \\
@@
     &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
-    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
-    &\qquad \qquad \qquad W' = [k \mapsto \iota_{b',e'}^0, i \mapsto \iota_{\malloc}]\\
-    &\qquad \qquad \qquad W' \future W \land \\
+    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot'\uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
     &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
+    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
     &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
@@
   \end{align*}
+
+  Additionally, we require that for every $W' \future [i\mapsto \iota_{\malloc,0}]$, $\heapSat[\hs]{n}{W'}$ implies that $\dom(\hs)$ contains $[n',+\infty)$ for some $n'$.
 \end{specification}

### 2016-11-18 local-capability-machine.tex
Started a new file with the local capabilities

--- 2016-11-18 capability-machine.tex
@@
 \newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
+\newcommand{\adv}{\var{adv}}
+\newcommand{\flag}{\var{flag}}
 
@@
 \Mems  &::= \Regs \times \Heaps \\
-\Caps  &::= (\Perms \times \Globals) \times \Addrs \times \Addrs \times \Addrs & & &
+\Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs & & &
 \Confs &::= \Mems + \{\failed, \halted \times \Mems\} \\
@@
 
+\section{Macros}
+In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
+\begin{description}
+\item[\texttt{call f(r)}] Moves the argument to register $\texttt{r}$ (if there are more arguments, then they are moved to the appropriate registers).
+Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return).
+Erases all registers but the argument register(s), $\pcreg$, and the return register.
+Jumps to $\texttt{f}$.
+\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory. In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
+\item[\texttt{hstore x n}] Fetches the capability refered to with \texttt{x} and stores \texttt{n} to it.
+\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them.
+If they are equal, then execution continues.
+If they are unequal, then the assertion flag is set to 1 and execution halts.
+\end{description}
+
 \section{Examples}
@@
 \subsubsection{Simple use of local capabilities}
+ML-like program:
+\begin{verbatim}
+let f = fun adv =>
+          let l = 1 in
+          adv();
+          assert (l == 1)
+\end{verbatim}
+Assembly program not using stack
+\begin{verbatim}
+f:  malloc l
+    hstore l 1
+    call adv()
+    assert l 1
+\end{verbatim}
+
 \begin{verbatim}
@@
 Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
-\lars{So this means that the value returned by *l is 1, correct ? We
-  could add an assert statement to the example to show it more directly.}
-\lau{Yes, and done.}
-
-\lars{What about an even simpler example, such as the one below:}
-\begin{verbatim}
-let f = fun adv =>
-          let l = 1 in
-          adv();
-          assert (l == 1)
-\end{verbatim}
-\lau{This example should certainly also hold true work, but it does not need local capabilities (it might serve well as a sanity check). }
+
+Correctness lemma not assuming stack
+\begin{lemma}
+  For all $n$,
+  let
+  \begin{align*}
+    c_\adv & = ((\entry,\local),\start_\adv,\addrend_\adv,\start_\adv) \\
+    c_f & = ((\exec,\local),\start_f,\addrend_f,\start_f)
+  \end{align*}
+  if
+  \[
+    (\reg_0\update{\pcreg}{c_f}\update{r_1}{c_\adv},\hs_\adv \uplus \hs_f \uplus \hs_\malloc \uplus h_\flag \uplus h_\var{frame}) \step[n] (\halted, h')
+  \]
+  where
+\lau{TODO: finish this}
+\end{lemma}
 
@@
 \begin{lemma}\forcenewline
-  \newcommand{\adv}{\var{adv}}
-  \newcommand{\flag}{\var{flag}}
   For all $n \in \nats$
@@
 
-\section{Old TD example}
-\subsection{Ticket Dispenser}
-\label{sec:tick-disp}
-\newcommand{\hsfoot}{\hs_\var{footprint}}
-\newcommand{\hsframe}{\hs_\var{frame}}
-\newcommand{\size}{\var{size}}
-\newcommand{\rio}{r_{io}}
-\newcommand{\adv}{\codelabel{adv}}
-\newcommand{\advb}{\var{adv_{base}}}
-\newcommand{\adve}{\var{adv_{end}}}
-\newcommand{\initb}{\var{init}_{base}}
-\newcommand{\inite}{\var{init}_{end}}
-\newcommand{\mrlen}{5cm}
-\newcommand{\retm}{\var{ret}_{\malloc}}
-\newcommand{\reta}{\var{ret}_{\adv}}
-\newcommand{\base}{\var{base}}
-\newcommand{\eend}{\var{end}}
-\newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
- \left . \vphantom{\begin{array}{l}
- \ifthenelse{\equal{#1}{1}}{3\\}{
-    \ifthenelse{\equal{#1}{2}}{3\\3\\}{
-    \ifthenelse{\equal{#1}{3}}{3\\3\\3\\}{
-    \ifthenelse{\equal{#1}{4}}{3\\3\\3\\3\\}{
-    \ifthenelse{\equal{#1}{5}}{3\\3\\3\\3\\3\\}{
-    \ifthenelse{\equal{#1}{6}}{3\\3\\3\\3\\3\\3\\}{
-      3\\3\\3\\3\\3\\3\\3\\ %7
-  }}}}}}
-  \end{array}} \right \}}}
-}
-\newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
-\lau{Has not been adjusted since local capabilities were added.}
-
-Assume the instructions of some adversary resides in the memory starting at the memory location marked with $\advb$.
-Assume that the register $\rio$ initially contains a capability for the address in memory where I/O is written to (we assume this address is 0).
-We assume that entry capabilities for $\adv$ and $\malloc$ is ambiently available\dominique{what does this mean?}.
-The following is a test program for a ticket dispenser:
-\[
-  \begin{array}{r l l p{\mrlen}}
-% Set io to -1
-\initb:     & \store{\refheap{\rio}}{-1} & \bracket{1} & \annotate{1}{Initialise io to $-1$} \\
-% Store io capability just after jump to adv
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store the io capability on the stack} \\
-           & \lea{\refreg{r_0}}{\iocap} & & \\
-           & \store{\refheap{r_0}}{\refreg{\rio}} & & \\
-% Overwrite io register
-           & \move{\refreg{\rio}}{0} & \bracket{1} & \annotate{1}{Overwrite io register} \\
-% Allocate memory for ticket dispenser program
-           & \move{\refreg{r_1}}{\size} &  \bracket{4} & \annotate{4}{Allocate memory for the ticket dispenser program (including memory for counter and capability for counter)}\\
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & & \\
-           & \lea{\refreg{r_0}}{3} & & \\
-           & \jmp{\malloc} & & \\
-% Save ticket dispenser code
-\retm:     & \store{\refheap{r_1}}{(\encode(i_1))} & \bracket{7} & \annotate{7}{Store the ticket dispenser program in the newly allocated memory} \\
-           & \lea{\refreg{r_1}}{1} & & \\
-           & \store{\refheap{r_1}}{(\encode(i_2))} & & \\
-           & \lea{\refreg{r_1}}{1} & & \\
-           & \vdots & & \\
-           & \store{\refheap{r_1}}{(\encode(i_7))} & & \\
-           & \lea{\refreg{r_1}}{1} & & \\
-% Save capability for counter
-           & \move{\refreg{r_0}}{\refreg{r_1}} & \bracket{4} & \annotate{4}{Store a capability for the counter address after the ticket dispenser program} \\
-           & \lea{\refreg{r_0}}{1} & & \\
-           & \store{\refheap{r_1}}{\refreg{r_0}} & & \\
-           & \lea{\refreg{r_1}}{1} & & \\
-% Save capability for counter two addresses after jump to adv.
-           & \move{\refreg{r_2}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Store a capability for the counter in this code} \\
-           & \lea{\refreg{r_2}}{\counter} & & \\
-           & \store{\refheap{r_2}}{r_0} & & \\
-           & \move{\refreg{r_2}}{0} & \bracket{1} & \annotate{1}{Overwrite capability for counter} \\
-% Save counter
-           & \store{\refheap{r_1}}{0} & \bracket{1} & \annotate{1}{Initialise counter to 0} \\
-% Change pointer to point to start of ticket dispenser code
-           & \lea{\refreg{r_1}}{-8} & \bracket{1} & \annotate{1}{Capability points to start of td code \footnote{This is also the argument for the adversary}} \\
-% Set up entry pointer to ticketDispenser for adv.
-           & \restrict{\refreg{r_1}}{\refreg{r_1}}{(\encodePerm{(\entry)})} &\bracket{1} & \annotate{1}{Restrict cap. to td} \\
-% Set up return pointer
-           & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Setup a return pointer for the adversary} \\
-           & \lea{\refreg{r_0}}{5} & & \\
-           & \restrict{\refreg{r_0}}{\refreg{r_0}}{(\encodePerm{(\entry)})} & & \\
-% jump to adv
-           & \jmp{\adv} & \bracket{1} & \annotate{1}{Jump to the adversary} \\
-           & 0 & \bracket{1} & \annotate{1}{Address reserved for io cap.} \\
-           & 0 & \bracket{1} & \annotate{1}{Address reserved for counter cap.} \\
-% retrieve io capability
-\reta:     & \move{\refreg{r_0}}{\refreg{\pcreg}} & \bracket{3} & \annotate{3}{Retrieve the io capability} \\
-           & \lea{\refreg{r_0}}{$-2$} & & \\
-           & \load{\refreg{\rio}}{\refheap{r_0}} & & \\
-% retieve counter capability
-           & \lea{\refreg{r_0}}{1} & \bracket{2} & \annotate{2}{Retrieve the counter capability} \\
-           & \load{\refreg{r_1}}{\refheap{r_0}} & & \\
-% load the value of the counter
-           & \load{\refreg{r_0}}{\refheap{r_1}} & \bracket{2} & \annotate{2}{Read the counter and write it to io address} \\
-% store the value of the counter to io
-           & \store{\refheap{\rio}}{\refreg{r_0}} & & \\
-\inite :   & \halt
-  \end{array}
-\]
-Here $\size$ is 9.
-The variables $\counter$ and $\iocap$ are respectively the offsets to the addresses reserved for the counter and io capabilities.
-$i_1 \dots i_7$ refers to the instructions in the following ticket dispenser:
-\[
-  \begin{array}{r l}
-    i_1 :& \move{\refreg{r_2}}{\refreg{\pcreg}} \\
-    i_2 :& \load{\refreg{r_2}}{\size - 2} \\
-    i_3 :& \load{\refreg{r_1}}{\refheap{r_2}}\\
-    i_4 :& \plus{\refreg{r_3}}{\refreg{r_1}}{2} \\
-    i_5 :& \store{\refheap{r_2}}{\refreg{r_3}} \\
-    i_6 :& \move{\refreg{r_2}}{0} \\
-    i_7 :& \jmp{\refreg{r_0}}
-  \end{array}
-\]
-The heap layout for the above program can be seen in Figure~\ref{tab:tick-disp-heap}.
-\begin{figure}[ht]
-  \centering
-  \begin{tabular}{ | c | l | }
-    \hline
-    Addr. & Content \\ \hline
-    0   &  $i_1$ \\ \hline
-    1   &  $i_2$ \\ \hline
-    2   &  $i_3$ \\ \hline
-    3   &  $i_4$ \\ \hline
-    4   &  $i_5$ \\ \hline
-    5   &  $i_6$ \\ \hline
-    5   &  $i_7$ \\ \hline
-    7   & Cap. for addr. 8 \\ \hline
-    8   & The counter \\ \hline  
-  \end{tabular}
-  \caption{The heap layout for the ticket dispenser.}
-  \label{tab:tick-disp-heap}
-\end{figure}
-
-In the JS paper, the ticket dispenser program dereferences the counter and returns the value.
-It does not seem like a similar thing would work here as the adversary can do even more.
-In particular, the adversary can execute a $\halt$ instruction that will cause the machine to halt (successfully).
-The adversary could also choose never to return to the return address that we set up for him.
-\begin{lemma}[Ticket dispenser test]\dominique{I would prefer more precise, symbolic statements of assumptions and less prose.}
-\label{lem:tckt-disp}
- Given a configuration $\Phi$, if
- \begin{description}
- \item[$\boldsymbol \memheap$] contains the ticket dispenser test program starting at the address labelled $\initb$, 
- \item[$\boldsymbol \memheap$] contains the adversary code starting at label $\advb$.
-The adversary is assumed to have no capabilities on the heap, but it has access to the ambient $\malloc$ capability!
-\lau{maybe mention that malloc is also on the heap}
- \item[$\boldsymbol \memheap$] contains the malloc code, and the malloc code satisfy $\iota_{\malloc}$, i.e., $\forall n \ldotp \heapSat[{\heap\restrictfun{\{\malloc_{\var{base}}, \dots ,\malloc_{\var{end}}\}}}]{n}{W_{\malloc}}$.
- \item[All the above programs are disjoint.]
- \item[$\boldsymbol{\adv}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
- \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\initb,\inite,\initb)$.
- \item[$\boldsymbol{ \memreg(\rio)}$] is $(\readwrite,0,0,0)$.
- \item[the remaining registers] all contains 0,
- \end{description}
- and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.
-\end{lemma}
-In the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.
-
-The ticket dispenser test program uses $\malloc$.
-To be able to reason about programs that use $\malloc$, we assume that it follows the following specification.
-If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory.
-\begin{specification}[Malloc v.2]
-  $c_\malloc$ satisfies the specification for malloc iff
-  \begin{align*}
-    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
-    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
-    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
-    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
-%    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\ % This
-    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
-    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
-    &\qquad \quad \memreg(r_1) = \size \land \\
-    &\qquad \quad \size \geq 0 \land \\
-    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
-    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
-    &\qquad \quad \Rightarrow \\
-    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
-    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
-    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
-    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot'\uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
-    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
-    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
-    &\qquad \qquad \qquad \size = e'-b' \\
-    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
-    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
-  \end{align*}
-
-  Additionally, we require that for every $W' \future [i\mapsto \iota_{\malloc,0}]$, $\heapSat[\hs]{n}{W'}$ implies that $\dom(\hs)$ contains $[n',+\infty)$ for some $n'$.
-\end{specification}
-The specification uses the malloc island defined as
-\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depend on the implementation (this is mostly a note to myself.)}
-\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
-  pc unspecified as it is not important for malloc users?}
-\dominique{Resulting frame should be equal to original.}
-\lau{Rewrite the specification when we have settled on a reasonable way to handle ABI.}
-
-\begin{specification}[Malloc]\dominique{can you state this more systematically, with less prose, more explicit assumptions and conclusion?}
-Take $W_{\malloc}$ the world with exactly one island, namely the unspecified island $\iota_{\malloc}$ that governs the internal state of $\malloc$.
-If we invoke $\malloc$ in a memory $h = \hsframe \uplus \hsfoot$
-%\lau{Is $\hsframe$ supposed to be the rest of memory, so $h$ is something we can execute in?
-Is $\malloc$ allowed to allocate more memory, i.e., how does $\hsfoot$ relate to $\hsfoot'$?} A: The frame contains anything that has been allocated and the footprint contains the rest.
-That is the footprint contains mallocs internal state as well as the free memory.
-% Notice: the current worlds do not support malloc allocating memory as the invariants cannot evolve.
-that is the disjoint sum of a frame part and a footprint part and the latter satisfies this world up to steps $n$, i.e., $\heapSat[\hsfoot]{n}{W_{\malloc}}$.
-Then $\malloc$ will successfully terminate and return a result in a heap $h' = \hsframe \uplus \hsfoot'$ that is the disjoint sum of the same unmodified frame part and a potentially new footprint part, such that there is a future world $W'$ of $W_{\malloc}$ consisting of exactly two islands, where the first is an instance of $\iota_{\malloc}$ and the second is the region $\iota_{\var{base},\var{end}}^0$
-%\lau{doesn't this invariant allow the newly allocated memory to contain capabilities?} A: It does not capture this. One possiblity is to make new standard all zero invariant.
- for an appropriately-sized interval $\var{base},\var{end}$ and the new footprint part satisfies this new world $\heapSat[\hsfoot']{n}{W'}$. 
-%\lau{at what step should this hold? Just some step $n'<n$?} A: try with n, if that does not work, then say something like: there exists j such that malloc terminates in j steps and use n-j.
-%\lau{It should probably be more precise what it means to ``invoke'' $\malloc$ and for $\malloc$ to ``return'' something.} A: yes.
-If we won't abstract notion of this, the standard way to formulate it would be with Hoare triples.
-%\lau{How do we express that $\malloc$ can be ``trusted'' in the sense that it won't save any capabilities like one to the newly allocated memory or the return capability.} A:  it allows for read capabilities!
-To capture this, we need either non-interference or some kind of trace semantics!
-
-To \emph{invoke} $\malloc$ means that the capability in $\pcreg$ points to a $\jmp{\malloc}$ instruction (and that this address is within its capabilities).
-Further in the register $r_0$ there has to be a return capability, i.e., a capability of the form $(\perm,\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm$ is either, $\rwx$, $\exec$, or $\entry$.
-Lastly, the register $r_1$ has to contain the argument for $\malloc$ which is the desired length of the piece of memory to be allocated.
-
-To \emph{return} from $\malloc$ means that the register $\pcreg$ contains the capability $(\perm',\var{returnStart},\var{returnEnd},\var{returnAddr})$ where $\perm'$ is $\rwx$ if $\perm$ was $\rwx$ and $\exec$ otherwise.
-Further, register $r_1$ contains $(\rwx,\var{base},\var{end},\var{base})$ which is a capability for the allocated memory.
-\end{specification}
-
- \begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
-Let configuration $\Phi$ be given where the heap is shaped as follows:
-\begin{itemize}
-\item $\memheap$ contains the ticket dispenser test program (tdtp) starting at address $\initb$.
-\item $\memheap$ contains the adversary code starting at address $\advb$ and ending at address $\adve$. The adversary code contains no capabilities.
-\item $\memheap$ contains the malloc code.
-\item (and so on) %TODO write the remaining assumptions
-\end{itemize}
-and the register file contains:
-\begin{enumerate}
-\item $\memreg(\rio) = (\readwrite,0,0,0)$ 
-\item $\memreg(\pcreg) = (\rwx,\initb, \inite, \initb)$
-\item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
-\end{enumerate}
-
-If we start running in $\Phi$ until $\jmp{\malloc}$ has just been executed, then the configuration is the same as $\Phi$ except:
-\begin{itemize}
-\item $\memheap(0)$ contains $-1$
-\item the io capability is saved to the designated address in the tdtp
-\item $r_1$ contains the size of the ticket dispenser program (9)
-\item $r_0$ contains the return capability $(\rwx,\initb,\inite, \retm)$
-\end{itemize}
-Call this new heap $\Phi'$
-
-To be able to say something about the jump to $\malloc$, we use its specification.
-Take $\hsfoot$ to be the part of the heap $\malloc$ resides in and everything that has not been allocated according to the assumptions.
-Take $\hsframe$ to be everything else.
-We have $\heapSat[\hsfoot]{n}{W_\var{\malloc}}$ as the world only has one region, namely the one that governs $\malloc$.\dominique{explicit assumption about this needed?} According to the specification of $\malloc$, we get $\Phi' \step^* \Phi''$ where
-\begin{itemize}
-\item $\memheap[\Phi'] = \hsframe \uplus \hsfoot'$
-\item $\memreg[\Phi'](r_1) = (\rwx,\var{base},\var{end},\var{base})$ for some addresses $\var{base}$ and $\var{end}$
-\item $\memreg[\Phi'](\pcreg) = (\rwx,\initb,\inite,\retm)$
-\end{itemize}
-Further for $W'$ with exactly two regions $\iota_\malloc$ and $\iota_{\var{base},\var{end}}^0$ we have that the new footprint satisfies this world, i.e., $\heapSat[\hsfoot']{n}{W'}$.
-
-If we continue running from $\Phi'$ until just after the point we have executed $\jmp{\adv}$, then  $\pcreg$ points to an execute capability for the adversary, and  we reach a configuration $\Phi''$ which is the same as $\Phi$ except:
-\begin{itemize}
-\item $\memheap[\Phi''](0)$ contains $-1$
-\item capabilities for \emph{the counter} and io are saved to the designated addresses in the tdtp
-\item the ticket dispenser code resides on the heap starting at the address $\var{base}$ (the code is as seen in Figure~\ref{tab:tick-disp-heap})
-\item $\memreg[\Phi''](\pcreg) = (\exec,\advb,\adve,\advb)$
-\item $\memreg[\Phi''](r_0) = (\entry,\initb,\inite,\reta)$
-\item $\memreg[\Phi''](r_1) = (\entry,\var{base},\var{end},\var{base})$
-\item all other registers contain 0
-\end{itemize}
-We now want to use the FTLR. To do so, we need to define a world that represents our current heap. We pick the following world:
-\begin{itemize}
-\item $\iota_{\var{io}}$ the region governs the heap segment that only has address $0$. The value of address $0$ should be $-1$ or $\geq 0$ and even.  % The region that governs that output address
-\item $\iota_{\var{td}}$ this region governs heap segments with the domain $[\var{base},\var{end}]$ for the addresses $[\var{base},\var{end} - 1]$ they are the same as in $\Phi''$.
-$\hs(\var{base})$ is positive and even.
-\item $\iota_{\var{tdtp}}$ this region governs the ticket dispenser test programs.
-The domain of a heap segment in this region is that of the ticket dispenser test program and the contents of the heap segment should be the same as it is in $\Phi''$.
-\item $\iota_{\advb,\adve}$
-\item $\iota_{\malloc}$
-\end{itemize}
-To use the FTLR, we have to argue that the read condition is satisfied for the region, the adversary is in.
-This is easily shown, as we have chosen the region for the adversary to be the standard one, so with $\advb$ and $\adve$ as witnesses, we need to show $\iota_{\advb,\adve} \subseteq \iota_{\advb,\adve}$.
-
-So we can conclude that $\npair{(\exec,\advb,\adve,\advb}) \in \stder(W'')$ where $W''$ is the world with the following regions:
-Assuming we can show the following three things:
-\begin{enumerate}
-\item $\memreg[\Phi''] \in \stdrr(W'')$
-\item $\npair{c = (\exec,\initb,\inite,\reta)}] \in \stdkr(W'')$
-\item $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$ where the we have limited the heap to the segments governed by the regions in $W''$.
-\end{enumerate}
-then it follows from the fact that the adversary program is in the expression predicate that
-\[
-\npair{\Phi^{(3)} = (\updateReg[\Phi'']{r_0}{c}[\plainproj{reg}.r_1 \mapsto (\exec,\advb,\adve,\advb)].\plainproj{reg},\memheap[\Phi''])} \in \observations(W'').
-\]
-From this and the fact that we know the execution terminates, we know that $\Phi^{(3)} \step^k (\halted,h)$ for some heap $h$ and $k \leq n$.
-From this we can conclude that the heap segment that only consists of address 0 is accepted by region $\iota_{\var{io}}$ which means that either $h(0)$ is $-1$ or positive and even which is exactly what the lemma says.
-
-It remains to show that 1-3 holds.
-For 1, all registers but $r_0$ and $r_1$ are easy as they contain $0$ which is always in the value predicate.
-For $r_0$ and $r_1$, we do need to do some reasoning.
-Both capabilities have one things in common, namely that the underlying programs do not use the arguments.
-
-$\memreg[\Phi''](r_0)$ contains the capability $(\entry,\initb,\inite,\reta)$ and we need to show that it is in $\stdvr(W'')$.
-Let $n' < n$ and $W^{(3)} \future W''$ be given and show $\npair[n']{(\exec,\initb,\inite,\reta)} \in \stder(W^{(3)})$.
-Let $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
-We do not really care about the register file and the continuation as the ticket dispenser test program does not use them.
-As $\heap$ satisfies $W^{(3)}$, parts of the heap must satisfy the invariants $\iota_{\var{tdtp}}$ and $\iota_{\var{td}}$ which means the ticket dispenser program and the ticket dispenser test program is unaltered.
-It also means that the counter in the ticket dispenser is $\geq 0$ and even.
-We need to show $\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\initb,\inite,\reta), \heap \uplus \heap_f)} \in \observations(W^{(3)})$.
-If $n'$ is sufficiently small, then we run out of steps before the program halts in which case the configuration is trivially in the $\observations$ predicate.
-If the execution halts successfully, then the ticket dispenser test program will have moved the value of the counter in the ticket dispenser to address 0, the designated I/O address.
-As this is a numerical value, it will trivially be in the value predicate.
-Further, the island that governs address 0 is $\iota_{\var{io}}$, so we have to argue that the value is $\geq 0$ and even.
-The value came from the ticket dispenser which is governed by the region $\iota_{\var{td}}$ so it is indeed $\geq 0$ and even.
-
-In register $r_1$ we have the capability $(\entry,\var{base},\var{end},\var{base})$ and we need to show $\npair[n]{(\exec,\var{base},\var{end},\var{base})} \in \stder(W'')$.
-To this end, we go through the same motion as for $r_0$:  Let $n' < n$, $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\npair[n']{c} \in \stdkr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given.
-Show
-\[
-\npair[n']{(\reg[r_0 \mapsto c][r_1 \mapsto ](\exec,\var{base},\var{end},\var{base}), \heap \uplus \heap_f)} \in \observations(W^{(3)}).
-\]
-Again the heap satisfaction makes sure that the ticket dispenser program is as we expect, so we can execute it.
-If we run out of steps before we reach $\jmp{\refreg{r_0}}$, then it is in the observation predicate.
-Otherwise, we reach a new configuration $\Phi^{(3)}$ where the registers are the same as in $\reg$ except for $r_0$, $r_1$, $r_2$, and $\pcreg$ which will be $c$, the previous counter value, $0$, and $\updatePcPerm(c)$ respectively.
-$0$ is trivially in the value predicate and $c$ is in the value predicate as it is in the continuation predicate, so $\reg$ is in the register-file predicate.
-If we split the heap part of $\Phi^{(3)}$ into the same two parts as we had in the initial assumption, then we need to argue that the non-frame part still satisfies the world.
-The only change made to the heap was updating the counter.
-The heap segments that did not change still satisfy their regions.
-We have to argue that the heap segment $[\var{base},\var{end}]$ still satisfies the region $\iota_{\var{td}}$.
-The ticket dispenser code remains unchanged and the counter is $\geq 0$ and even as it was computed by adding two to a value which was $\geq 0$ and even, so the ticket dispenser heap segment satisfies $\iota_{\var{td}}$.
-Now we have all the necessary conditions satisfied to be able to use the fact that $c$ is in the continuation predicate to conclude
-
-\lau{did I mess up the steps here?}
-\[
-\npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\updatePcPerm{c}}, \heap' \uplus \heap_f)} \in \observations
-\]
-(where $\heap'$ is the new heap). From this we get that continuing from the configuration we end up with no more steps, failing, or in a halting state that satisfies the world.
-
-%condition 2
-
-\dominique{this duplicates the argument about $(\entry,\initb,\inite,\initb+\var{ret}) \in \stdvr(W'')$. Find a way to share this?}
-For condition 2, we need to show 
-\[
-\npair{c = (\entry, \initb, \inite,\reta)}] \in \stdkr(W'').
-\]
-To this end, let $W^{(3)} \future W''$, $\reg \in \stdrr(W^{(3)})$, $\heapSat[\heap]{n'}{W^{(3)}}$ and $\heap_f$ be given. We then need to argue 
-\[
-\npair{(\reg\update{\pcreg}{\updatePcPerm{c}},\heap \uplus \heap_f)} \in \observations(W^{(3)}).
-\]
-As part of the heap satisfies a future world of $W''$, we know that the ticket dispenser test program remains unaltered.
-If we run it (and do not run out of steps), then it loads the counter, stores it to address $0$, and halts.
-From the heap satisfaction, we get that the counter is positive and even, so the value we store to address 0 is positive and even.
-The value is trivially in the value relation as it is an integer.
-Further the region that governs address 0 is $\iota_{\var{io}}$ which requires the value there to be positive and even which it is.
-So in this case the configuration is in the observation predicate.
-If we run out of steps, then it is also in the observation relation.
-
-%condition 3
-Finally, we need to show $\heapSat[{\memheap[\Phi'']}|]{n}{W''}$.
-The domains of the heap segments are assumed to be disjoint, and we have limited the heap to only be the parts considered in the regions, so the first condition of heap satisfaction is okay.
-Let us take the regions one at a time and briefly argue that the corresponding heap segment satisfies the region:
-\begin{itemize}
-\item $\iota_{\var{io}}$ governs address 0, which i initial $-1$. This region requires it to be $-1$ or $\geq 0$ and even, so it is initially satisfied.
-\item $\iota_{\var{tdtp}}$, this region was constructed so the part of the heap it governs remains unaltered.
-\item $\iota_{\var{td}}$, apart from the address with the counter, this region is also constructed so this heap segment does not change.
-With respect to the counter, it requires it to be $\geq 0$ and even.
-The counter starts as 0, so also this cell satisfies the region.
-\item $\iota_{\advb,\adve}$ the adversary code is assumed to be instructions only which are trivially satisfied by the standard island.
-\item $\iota_{\malloc}$, from the specification of $\malloc$ we got that the resulting heap after the allocation satisfied a world with $\iota_{\malloc}$.
-We use the same region and we have not changed that part of the heap, so it still satisfies this region.
-\end{itemize}
-
-We have now shown the three conditions and thus proven Lemma~\ref{lem:tckt-disp}.
-\end{proof}
-
-\section{Logical Relation}
-\label{sec:logical-relation}
-\subsection{Recursive Domain Equation}
-\label{subsec:recursive-dom-eq}
-The goal is to solve the following domain equation:
-\[
-\Wor = \nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
-\]
-Where $\States$ is a set of states with all the ones we use in this paper.
-\[
-\Rel= \{R \in \powerset{\States^2} \mid R \text{ is reflexive and transitive} \}
-\]
-
-This cannot be solved with sets, so we use preordered complete ordered families of equivalences where it is possible to solve such an equation that resembles the above one, namely it is possible to find an isomorphism $\xi$ and preordered \cofe{} $W$ such that
-\[
-  \xi : \Wor \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
-\]
-
-\begin{definition}[o.f.e's]
-  An ordered family of equivalences (o.f.e.) is a set and a family of equivalences, $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$.
-The family of equivalences have to satisfy the following properties
-  \begin{itemize}
-  \item $\nequal[0]$ is a total relation on $X$
-  \item $\forall n \ldotp \forall x,y \in S \ldotp x \nequal[n+1] y \Rightarrow x \nequal y$
-  \item $\forall x,y \ldotp(\forall n \ldotp x \nequal y) \Rightarrow x=y$
-  \end{itemize}
-\end{definition}
-
-\dominique{I suppose you're using a standard ultrametric metric to make an
-  o.f.e. a metric space?}
-\begin{definition}[\cofes]
-  A complete ordered family of equivalences is an o.f.e.\ $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ where all Cauchy sequences in $X$ have a limit in $X$.
-\end{definition}
-
-\begin{definition}[Preordered \cofes]
-  A preordered \cofe{}\ is a \cofe{}\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
-  \begin{itemize}
-  \item The ordering preserves limits. That is, for Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
-  \end{itemize}
-  % Preorder:
-  % \begin{itemize}
-  % \item $\forall x \in X \ldotp x \future x$
-  % \item $\forall x,y,z \in X \ldotp x \future y \land y \future z \Rightarrow x \future z$
-  % \end{itemize}
-\end{definition}
-
-\begin{definition}[Preordered \cofe{}\ construction: Finite-partial function]
-  Given a set $S$ and preordered \cofe{}\ $X$, $S \finparfun X$ is a preordered \cofe{}\ with the ordering
-  \[
-    \begin{gathered}
-      f \future g  \\
-      \text{ iff } \\
-      \dom(f) \supseteq \dom(g) \; \text{ and } \; \forall n \in S \ldotp f(n) \future g(n)
-    \end{gathered}
-  \]
-\end{definition}
-
-We need the following constructions to create the preordered \cofe{} needed to solve the recursive domain equation.
-\begin{definition}[Preordered \cofe{}\ construction: Function]
-  Given a set $S$ and \cofe{}\ $\var{HP}$, $S \fun \var{HP}$ is a preordered \cofe{}\ with the ordering
-\[
-  \begin{gathered}
-    f \future g \\
-    \text{ iff } \\
-    \forall s \in \dom(g) \ldotp f(s) \future g(s)
-  \end{gathered}
-\]
-\end{definition}
-
-\begin{definition}[Preordered \cofe{}\ construction: Monotone, non-expansive function]
-  Given a preordered \cofe{}\ $W$ and preordered \cofe{}\ $U$, $W \monnefun U$ is a preordered \cofe{} with the ordering
-\[
-  \begin{gathered}
-    f \future g \\
-    \text{ iff } \\
-    \forall s \in \dom(g) \ldotp f(s) \future g(s) 
-  \end{gathered}
-\]
-\end{definition}
-
-The above are standard constructions, so they are used here without showing they
-are in fact well-defined as shown in \cite{Birkedal:tutorial-notes}.
-
-\begin{definition}[Preordered \cofe{}\ construction: Region]
-  \label{def:pcofe-region}
-  Given a \cofe{}\ $H$, the tuple 
-\[
-  (\States \times \Rel \times H)
-\]
- is a preordered \cofe{}\ with the ordering
-\[
-  \begin{gathered}
-    (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_2)\\
-    \text{ iff }\\
-    H_2 = H_1 \text{ and } \phi_2 = \phi_1 \text{ and } (s_1,s_2) \in \phi_2\\
-  \end{gathered}
-\]
-\end{definition}
-\begin{lemma}[Region definition well-defined]
-\label{lem:reg-def-wd}
-  The construction in Definition~\ref{def:pcofe-region} is a preordered \cofe{}. That is
-  \begin{itemize}
-  \item It is a \cofe{} (this is a standard construction)
-  \item $\future$ is a transitive and reflexive relation.
-  \item $\future$ preserves limits. \\That is for Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$ if 
-    \[
-      \{a_n\}_n \geq \{b_n\}_n,
-    \]
-    then 
-    \[
-      \lim \{a_n\}_n \future \lim \{b_n\}_n
-    \]
-  \end{itemize}
-\end{lemma}
-
-The category of \cofes{} is the category with \cofes{} as objects and non-expansive functions as morphisms.
-We denote this category $\CatC$.
-The category of preordered \cofes{} has preordered \cofes{} as objects and monotone and non-expansive functions as morphisms.
-We denote this category $\CatP$.
-
-Define functors $K$, $R$, and $G$ as follows:
-\begin{align*}
-  & K : \CatP \rightarrow \CatP \\
-  & K(R) = \nats \finparfun R \\
-  & K(f) = \lambda \phi \ldotp \lambda n \ldotp f(\phi(n)) \\
-  & \\
-  & R : \CatC \rightarrow \CatP \\
-  & R(H) = \States \times \Rel \times H\\
-  & R(h) = \lambda (s,\Phi,H) \ldotp (s,\Phi,h(H))\\
-  & \\
-  & G : \CatP^{\var{op}} \rightarrow \CatC \\
-  & G(W) = \States \nefun W \monnefun \UPred{HS} \\
-  & G(g) = \lambda H \ldotp \lambda \var{st} \ldotp \lambda x \ldotp H (\var{st})(g(x))
-\end{align*}
-We first show that $K$, $R$, and $G$ are well-defined mappings.
-\begin{lemma}[World finite partial mapping]
-  \label{lem:world-finpar}
-  For all $f$ and $\phi$, $K(f)(\phi)$ is a finite partial mapping.
-\end{lemma}
-\begin{lemma}[Heap segment predicate monotone]
-\label{lem:HSP-mon}
-For all $g$, $H$, and $\var{st}$
-\[
-G(g) (H) (\var{st})
-\]
-is non-expansive.
-\end{lemma}
-\begin{lemma}[Heap segment predicate non-expansive]
-\label{lem:HSP-ne}
-For all $g$, $H$, and $\var{st}$
-\[
-G(g) (H) (\var{st})
-\]
-is monotone.
-\end{lemma}
-Next we show that $K$, $R$, and $G$ are in fact functors:
-\begin{lemma}[$K$ functorial]\forcenewline
-\label{lem:K-func}
-  \begin{enumerate}
-  \item $K(f) : K(X) \rightarrow K(Y)$ is monotone and non-expansive for $f : X \monnefun Y$
-  \item $K(f \circ g) = K(f) \circ K(g)$ for $f : Z \nefun Y$ and $g : X \nefun Z$
-  \item $K(\id) = \id$
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}[$R$ functorial]\forcenewline
-\label{lem:R-func}
-  \begin{enumerate}
-  \item $R(f) : R(X) \rightarrow R(Y)$ is non-expansive and monotone for $f : X \nefun Y$
-  \item $R(f \circ g) = R(f) \circ R(g)$ for $f : Z \nefun Y$ and $g : X \nefun Z$
-  \item $R(\id) = \id$
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}[$G$ functorial]\forcenewline
-\label{lem:G-func}
-  \begin{enumerate}
-  \item $G(f) : G(Y) \rightarrow G(X)$ is non-expansive for $f : X \monnefun Y$
-  \item $G(f \circ g) = G(g) \circ G(f)$ for $f : Z \nefun Y$ and $g : Y \nefun Z$
-  \item $G(\id) = \id$
-  \end{enumerate}
-\end{lemma}
-We now compose the above functors into the functor we actually want to use: $F = K \circ R \circ G$, $F : \CatP^{\var{op}} \rightarrow \CatP$.
-\begin{lemma}[$F$ functorial]\forcenewline
-\label{lem:F-func}
-  \begin{enumerate}
-  \item $F(f) : F(Y) \rightarrow F(X)$ is monotone and non-expansive for $f : X \monnefun Y$
-  \item $F(f \circ g) = F(g) \circ F(f)$ for $f : Z \nefun Y$ and $g : Y \nefun Z$\
-  \item $F(\id) = \id$
-  \end{enumerate}
-\end{lemma}
-
-\begin{lemma}[$F$ locally non-expansive]
-\label{lem:F-loc-ne}
-For all $f, g : X \fun Y$, if $f \nequal g$, then $F(f) \nequal F(g)$.
-\end{lemma}
-With $F$ being locally-non-expansive, we can pre- or post-compose with later ($\blater$) to get a locally contractive function. In this case we construct $F'$ by post-composition of $\blater$:
-\[
-  F'(\Wor) = \blater (F(\Wor))
-\]
-We have a theorem that gives us a solution to the recursive domain equation
-\[
-  \Wor \cong F'(\Wor) = \blater (\nats \finparfun (\States \times \Rel \times (\States \fun \Wor \monnefun \UPred{\HeapSegments} )))
-\]
-The solution to the recursive domain equations is presented by \cite{Birkedal:2010:TCS:411:4102-4122}.
-They solve it in pre-ordered, non-empty, complete, 1-bounded ultrametric spaces, but they have a simple correspondence to pre-ordered \cofes{}.
-
-\subsection{Worlds}
-\label{subsec:worlds}
-Assume preordered \cofe{}\ $\Wor$ and isomorphism $\xi$ such that:
-\[
-  \xi : \Wor \cong \blater (\nats \finparfun (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
-\]
-$\xi$ is a morphism in the category $\CatP$ (the category of preordered \cofes{}), so it is non-expansive and monotone.
-
-We now define regions as
-\[
-\Regions \defeq (\States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
-\]
-define region names to be natural numbers, i.e.,
-\[
-  \RegionNames \defeq \nats 
-\]
-and define worlds as
-\[
-\Worlds \defeq \RegionNames \finparfun \Regions
-\]
-To define future worlds and regions, We use the ordering inherited from the preordered \cofes{}.
-\begin{definition}[Future regions]
-For regions $(s_2,\phi_2,H_2), (s_1,\phi_1,H_1) \in \Regions$
-  \begin{align*}
- (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) &&\text{ iff } & &
-(\phi_1,H_1) = (\phi_2,H_2) \text{ and } (s_1,s_2) \in \phi_2
-  \end{align*}
-\end{definition}
-
-\begin{definition}[Future worlds]
-  For $W, W' \in \Worlds$
- \begin{align*}
- W' \future W & & \text{ iff } & &
-   \begin{gathered}
-     \dom(W') \supseteq \dom(W)\\ 
-     \text{ and } \\
-     \forall r \in \dom(W) \ldotp W'(r) \future W(r)
-   \end{gathered}
- \end{align*}
-\end{definition}
-
-\begin{lemma}[Future worlds transitive]
-\label{lem:future-worlds-trans}
-  \begin{align*}
-    & \forall W, W', W'' \ldotp \\
-    & \quad  W'' \future W' \land W' \future W \Rightarrow W'' \future W
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}
-\label{lem:nequal-and-future-world}
-  \begin{align*}
-    & \forall W_1,W_2,W_1' \ldotp \\
-    & \quad W_1 \nequal W_2 \land W_1' \future W_1 \Rightarrow \\
-    & \qquad \exists W_2' \ldotp W_2' \future W_2 \land W_1' \nequal W_2'
-  \end{align*}
-\end{lemma}
-
-\begin{definition}[$n$-subset for regions]
-For regions $(s_1,\phi_1,H_1), (s_2,\phi_2,H_2) \in \Regions$
-\begin{align*}
-  (s_1,\phi_1,H_1) \nsubeq (s_2,\phi_2,H_2) & & \text{ iff } & &
-  \begin{gathered}
-    (s_1,\phi_1) = (s_2,\phi_2)\\
-    \text{ and }\\
-    \forall W \in \Wor \ldotp H_1 \; s_1 \; W \nsubeq H_2 \; s_2 \; W
-  \end{gathered}
-\end{align*}
-\end{definition}
-
-\begin{definition}[Heap satisfaction/erasure]
-\begin{align*}
-&\heapSat[\hs]{n}{W} \\
-&\qquad\text{ iff }\\
-&\exists R : \dom(W) \rightarrow \HeapSegments \ldotp & &\\
-&\quad\begin{aligned}
-  &\hs = \biguplus_{r \in \dom(W)} R(r) \\
-  &\qquad\text{and}\\
-  &\forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\xi^{-1}(W))
-\end{aligned} 
-\end{align*}
-o\end{definition}
-
-\begin{lemma}[Heap satisfaction uniformity]
-\label{lem:heapsat-dc}
-  \begin{align*}
-    & \forall n,n',\hs, W \ldotp\\
-    & \quad \heapSat[\hs]{n}{W} \land n' \leq n \Rightarrow \heapSat[\hs]{n'}{W}
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Heap satisfaction non-expansive]
-\label{lem:heapsat-ne}
-  \begin{align*}
-    & \forall n,\hs, W, W' \ldotp\\
-    & \quad \heapSat[\hs]{n}{W} \land W \nequal W' \Rightarrow \heapSat[\hs]{n}{W'}
-  \end{align*}
-\end{lemma}
-
-\subsection{Logical Relation Setup}
-\label{subsec:logical-relation-setup}
-Our logical relation is defined using multiple recursive definitions, so the definitions in the following subsections are defined simultaneously.
-We want to define the value relation as the fixed-point given by Banach's fixed-point theorem, so all our definitions will be parameterised with the value relation.
-\subsubsection{Observation Relation}
-\label{subsubsec:observation-relation}
-In order to define the expression relation, we define an observation relation.
-\begin{align*}
-  \observations \quad : & \quad  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
-                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
-                           & \qquad \Rightarrow \exists W' \future W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
-\end{align*}
-A pair of a register and a heap segment is ``good'' if we can put it together with a frame heap, so we can execute it.
-The execution should then end up in a heap where the frame remains the same and the remaining heap segment satisfies the world.
-
-Note that the operational semantic is total, so we cannot get stuck.
-If the execution ends up in a $\failed$ configuration, then we do not care about the heap and the registers.
-This is why, we only have requirements on the result when we end up in a $\halted$ configuration.
-
-The following lemmas show that the observation relation is well-defined.
-\begin{lemma}[Observation relation uniformity]
-\label{lem:obs-dc}
-  \begin{align*}
-    &\forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp\\
-    &\quad \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Observation relation non-expansive in worlds]
-\label{lem:obs-ne-worlds}
-\begin{align*}
-  &\forall W, W', n \ldotp \\
-  &\quad W \nequal W' \Rightarrow \observations(W) \nequal \observations(W')
-\end{align*}
-\end{lemma}
-
-\subsubsection{Register-File Relation}
-\label{subsubsec:register-file-relation}
-This relation is used in the definition of the continuation relation as well as the expression relation.
-\begin{align*}
-  \stdrr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Regs} \\
-  \stdrr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp
-                     \begin{aligned}[t]
-                       \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
-                                            & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
-                     \end{aligned}
-\end{align*}
-Well-formedness lemmas for this definition:
-\begin{lemma}[Register relation uniformity]
-\label{lem:reg-dc}
-  \begin{align*}
-    &\forall \stdvr, n' \leq n \ldotp \forall W \ldotp \forall \reg \ldotp\\
-    &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair[n']{\reg} \in \stdrr(\stdvr)(W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Register relation montone in worlds]
-\label{lem:reg-mono-worlds}
-  \begin{align*}
-    &\forall \stdvr, n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp\\
-    &\quad \npair{\reg} \in \stdrr(\stdvr)(W) \Rightarrow \npair{\reg} \in \stdrr(\stdvr)(W')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Register relation non-expansive in worlds]
-  \label{lem:reg-ne-worlds}
-  \begin{align*}
-    &\forall \stdvr, n \ldotp \forall W, W' \ldotp \\
-    &\quad W \nequal W' \\
-    &\qquad \Rightarrow \stdrr(\stdvr)(W) \nequal \stdrr(\stdvr)(W')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Register relation non-expansive in value relation]
-\label{lem:reg-ne-vr}
-\begin{align*}
-  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stdrr(\stdvr) \nequal \stdrr(\stdvr')
-\end{align*}
-\end{lemma}
-
-\subsubsection{Continuation Relation}
-\label{subsubsec:continuation-relation}
-The continuation relation is used in the definition of the expression relation.
-The continuation relation ensures that if you continue execution through a continuation, then it will result in a good result according to the world.
-\begin{align*}
-  \stdkr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
-  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
-                     \begin{aligned}[t]
-                       & \npair{c} \in \stdvr(W) \land \\
-                       &  \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
-                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(\stdvr)(W') \ldotp \\
-                       & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
-                 \end{aligned}
-\end{align*}
-We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.
-
-Normally, it would just be required for the return value to be in the value relation, so the continuation is invoked with this value.
-Here everything that is in the register file is available when the continuation is invoked, so we do not special case on $r_1$ (which is the register that in the calling convention we use should contain the return value).
-
-Well-definedness lemmas:
-\begin{lemma}[Continuation relation uniformity]
-\label{lem:stdkr-dc}
-  \begin{align*}
-    &\forall \stdvr \ldotp \forall n' \leq n \ldotp \forall W \ldotp \forall c \ldotp\\
-    &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair[n']{c} \in \stdkr(\stdvr)(W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Continuation relation monotone in worlds]
-\label{lem:stdkr-mono-worlds}
-  \begin{align*}
-    &\forall \stdvr \ldotp \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp\\
-    &\quad \npair{c} \in \stdkr(\stdvr)(W) \Rightarrow \npair{c} \in \stdkr(\stdvr)(W')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Continuation relation non-expansive in worlds]
-  \label{lem:stdkr-ne-worlds}
-  \begin{align*}
-    &\forall \stdvr, n \ldotp \forall W, W' \ldotp\\
-    &\quad W \nequal W' \\
-    &\qquad \Rightarrow \stdkr(\stdvr)(W) \nequal \stdkr(\stdvr)(W')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Continuation relation non-expansive in value relation]
-\label{lem:stdkr-ne-vr}
-\begin{align*}
-  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stdkr(\stdvr) \nequal \stdkr(\stdvr')
-\end{align*}
-\end{lemma}
-
-\subsubsection{Expression Relation}
-\label{subsubsec:expression-relation}
-The expression relation is defined as follows:
-\begin{align*}
-  \stder \quad : & \quad  (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \nefun \UPred{\Words} \\%TODO: Check whether this needs to be monotone
-  \stder \defeq & \lambda \stdvr \ldotp \lambda W \ldotp \{ \npair{\pc} \mid 
-  \begin{aligned}[t]
-    & \forall n' \leq n \ldotp\\
-    & \quad \forall \npair[n']{\reg} \in \stdrr(\stdvr)(W) \ldotp \\
-    & \quad \quad \forall \npair[n']{c} \in \stdkr(\stdvr)(W) \ldotp \\
-    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
-    & \quad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
-  \end{aligned}
-\end{align*}
-Well-definedness lemmas:
-\begin{lemma}[Expression relation uniformity]
-\label{lem:stder-dc}
-  \begin{align*}
-    & \forall \stdvr \ldotp \forall n' \leq n \ldotp \forall W \ldotp \forall \pc \ldotp\\
-    & \quad \npair{\pc} \in \stder(\stdvr)(W) \Rightarrow \npair[n']{\pc} \in \stder(\stdvr)(W) 
- \end{align*}
-\end{lemma}
-
-\begin{lemma}[Expression relation non-expansive in world]
-\label{lem:stder-ne-worlds}
-\begin{align*}
-  & \forall \stdvr \ldotp \forall W_1 \nequal W_2 \ldotp \stder(\stdvr)(W_1) \nequal \stder(\stdvr)(W_2)
-\end{align*}
-\end{lemma}
-
-\begin{lemma}[Expression relation non-expansive in value relation]
-\label{lem:stder-ne-vr}
-\begin{align*}
-  &\forall \stdvr, \stdvr', n \ldotp \stdvr \nequal \stdvr' \Rightarrow \stder(\stdvr) \nequal \stder(\stdvr')
-\end{align*}
-\end{lemma}
-
-\subsubsection{Standard Region}
-\label{subsubsec:standard-region}
-The following standard region is used in the definition of the value relation. Specifically, it is used in the $\readCond{}$ and the $\writeCond{}$ (to be defined next)
-
-\begin{align*}
-  \iota_{\var{base},\var{end}} & : \; (\Worlds \monnefun \UPred{\Words}) \nefun \Regions\\
-  \iota_{\var{base},\var{end}} & \defeq \lambda \; \stdvr \ldotp ((\var{base},\var{end}),=,H_{\var{std}} (\stdvr)) \\ & \\
-  H_{\var{std}} & : (\Worlds \monnefun \UPred{\Words}) \nefun \States \nefun \Wor \monnefun \UPred{\HeapSegments} \\
-  H_{\var{std}} \; \stdvr \; (\var{base},\var{end})\; \hat{W} &  \defeq \left\{ \npair{\hs} \middle|
-                                               \begin{aligned}
-                                                 & \dom(\hs) = [\var{base},\var{end}] \land \\
-                                                 & \forall \addr \in [\var{base},\var{end}] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi \; \hat{W})
-                                               \end{aligned} \right\}
-\end{align*}
-As mentioned previously, the set of states contains the ``necessary'' states. For the above to make sense, the set of states contains pairs of natural numbers $(\var{base},\var{end})$.
-
-The well-definedness lemmas for the above is:
-\begin{lemma}[$H_{\var{std}}$ is monotone in the worlds]
-\label{lem:std-reg-mono-world}
-\begin{align*}
-  & \forall \stdvr \ldotp \forall \start, \addrend \ldotp \forall W' \future W \ldotp \\
-  & \quad H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W' \supseteq H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W
-\end{align*}
-\end{lemma}
-
-\begin{lemma}[$H_{\var{std}}$ is non-expansive in the worlds]
-\label{lem:std-reg-ne-world}
-\begin{align*}
-  & \forall \stdvr \ldotp \forall \start, \addrend \ldotp \forall n \ldotp \forall W_1 \nequal W_2 \ldotp \\
-  & \quad H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W_1 \nequal H_{\var{std}}\; \stdvr \; (\var{base}, \var{end}) \; W
-\end{align*}
-\end{lemma}
-
-\begin{lemma}[$H_{\var{std}}$ is non-expansive in the value relation]
-\label{lem:std-reg-ne-vr}
-\begin{align*}
-  & \forall \stdvr, \stdvr' \ldotp \forall n \ldotp \stdvr \nequal \stdvr' \Rightarrow\\
-  & \quad H_{\var{std}}\; \stdvr \nequal H_{\var{std}}\; \stdvr'
-\end{align*}
-\end{lemma}
-
-\begin{lemma}[$H_{\var{std}}$ is uniform]
-  \label{lem:std-reg-dc}
-  \begin{align*}
-    & \forall \stdvr, s=(\start,\addrend), \hat{W}, n', n, \hs \ldotp \\
-    & \quad n' \leq n \land \npair{\hs} \in H_{\var{std}} \; \stdvr \; s \; \hat{W} \\
-    & \qquad \Rightarrow \npair[n']{\hs} \in H_{\var{std}} \; \stdvr \; s \; \hat{W}
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[$H_{\var{std}}$ is non-expansive in state]
-  \label{lem:std-reg-ne-state}
-  \begin{align*}
-    & \forall \stdvr, (\start,\addrend), (\start',\addrend'), n \ldotp \\
-    & \quad  (\start,\addrend) \nequal (\start',\addrend') \\
-    & \qquad \Rightarrow H_{\var{std}} \; \stdvr \; (\start,\addrend) \nequal H_{\var{std}} \; \stdvr \; (\start',\addrend')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[$\iota_{\start,\addrend}$ is non-expansive in the value relation]
-\label{lem:std-reg-i-ne-vr}
-\begin{align*}
-  & \forall \start, \addrend \ldotp \forall \stdvr, \stdvr' \ldotp \forall n \ldotp \stdvr \nequal \stdvr' \Rightarrow\\
-  & \quad \iota_{\start,\addrend}\; \stdvr \nequal \iota_{\start,\addrend}\; \stdvr'
-\end{align*}
-\end{lemma}
-
-\subsubsection{Capability Conditions}
-\label{subsubsec:capability-conditions}
-The definition of the value relation has the same conditions several times, so to define it concisely, we define the following conditions.
-\begin{align*}
-  & \readCond{} : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
-  & \readCond{}(\stdvr)(\start,\addrend,W) =        
-  \begin{aligned}[t]
-    \{ n \mid & \;\exists r \in \RegionNames \ldotp \\
-              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-              & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'}(\stdvr) \}
-  \end{aligned}\\
- & \\
-  & \writeCond{} : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats} \\
-  & \writeCond{}(\stdvr)(\start,\addrend, W) = 
-  \begin{aligned}[t]
-    \{ n \mid & \; \exists r \in \RegionNames \ldotp \\
-              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-              & \; \qquad W(r) \nequal[n-1] \iota_{\start',\addrend'}(\stdvr)  \}
-  \end{aligned}\\
- & \\
-  & \execCond{} : \; (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^2 \times \Perms \times \Worlds) \monnefun \DCPred{\nats} \\
-  & \execCond{}(\stdvr)(\start,\addrend,\perm,W) = 
-  \begin{aligned}[t]
-    \{ n \mid & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-              & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
-              & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(\stdvr)(W') \}
-  \end{aligned} \\
- & \\
-  & \entryCond{} : \;  (\Worlds \monnefun \UPred{\Words}) \nefun (\Addrs^3 \times \Worlds) \monnefun \DCPred{\nats} \\
-  & \entryCond{}(\stdvr)(\start,\addrend,\addr,W) =
-  \begin{aligned}[t]
-    \{ n \mid & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
-              & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(\stdvr)(W') \}
-  \end{aligned}
-\end{align*}
-
-The following lemmas show that the above conditions are well-defined:
-\begin{lemma}[Read condition downwards-closed]
-\label{lem:rc-dc}
-  \begin{align*}
-    &\forall \stdvr, n,n',W,\start, \addrend \ldotp \\
-    &\quad  n \in \readCond{}(\stdvr)(\start,\addrend,W) \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow n' \in \readCond{}(\stdvr)(\start,\addrend,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Read condition monotone in world]
-\label{lem:rc-mono-world}
-  \begin{align*}
-    &\forall \stdvr, n,W,W',\start, \addrend, \start',\addrend' \ldotp \\
-    &\quad  (\start',\addrend',W') \future (\start,\addrend,W) \\
-    &\qquad \Rightarrow \readCond{}(\stdvr)(\start',\addrend',W') \supseteq \readCond{}(\stdvr)(\start,\addrend,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Read condition non-expansive in worlds]
-\label{lem:rc-ne-world}
-  \begin{align*}
-    & \forall \stdvr, \start, \addrend, \start', \addrend', n, W, W' \ldotp \\
-    & \quad (\start, \addrend,W) \nequal (\start', \addrend',W') \Rightarrow \\
-    & \qquad \readCond{}(\stdvr)(\start, \addrend,W) \nequal \readCond{}(\stdvr)(\start', \addrend',W')\\
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Read-write condition uniformity]
-\label{lem:wc-dc}
-  \begin{align*}
-    &\forall \stdvr, n,n',W,\start, \addrend \ldotp \\
-    &\quad  n \in \writeCond{}(\stdvr)(\start,\addrend,W) \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow n' \in \writeCond{}(\stdvr)(\start,\addrend,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Read-write condition monotone in world]
-\label{lem:wc-mono-world}
-  \begin{align*}
-    &\forall \stdvr,W,W',\start, \addrend, \start',\addrend' \ldotp \\
-    &\quad  (\start',\addrend',W') \future (\start,\addrend,W) \\
-    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start',\addrend',W') \supseteq \writeCond{}(\stdvr)(\start,\addrend,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Read-write condition non-expansive in world]
-  \label{lem:wc-ne-world}
-  \begin{align*}
-    &\forall \stdvr, n,W,W',\start, \addrend, \start',\addrend' \ldotp \\
-    &\quad  (\start',\addrend',W') \nequal (\start,\addrend,W) \\
-    &\qquad \Rightarrow \writeCond{}(\stdvr)(\start,\addrend,W) \nequal \writeCond{}(\stdvr)(\start',\addrend',W')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Execute condition downwards-closed]
-\label{lem:ec-dc}
-  \begin{align*}
-    &\forall \stdvr, n,n',W,\start, \addrend, \perm \ldotp \\
-    &\quad  n \in \execCond{}(\stdvr)(\start,\addrend,\perm,W) \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow n' \in \execCond{}(\stdvr)(\start,\addrend,\perm,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Execute condition monotone in world]
-\label{lem:ec-mono-world}
-  \begin{align*}
-    &\forall \stdvr,n,W, W',\start, \addrend, \perm,\start', \addrend', \perm' \ldotp \\
-    &\quad  (\start',\addrend',\perm',W') \future (\start,\addrend,\perm,W) \\
-    &\qquad \Rightarrow \execCond{}(\stdvr)(\start',\addrend',\perm',W') \supseteq \execCond{}(\stdvr)(\start,\addrend,\perm,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Execute condition non-expansive in worlds]
-\label{lem:ec-ne-world}
-  \begin{align*}
-    &\forall \stdvr, W, W',n,\start,\addrend,\perm,\start',\addrend',\perm' \ldotp \\
-    &\quad (\start,\addrend,\perm,W) \nequal (\start',\addrend',\perm',W') \Rightarrow \\
-    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\perm,W) \nequal \execCond{}(\stdvr)(\start',\addrend',\perm',W')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Entry condition downwards-closed]
-\label{lem:en-dc}
-  \begin{align*}
-    &\forall \stdvr, n,n',W,\start, \addrend, \addr \ldotp \\
-    &\quad  n \in \entryCond{}(\stdvr)(\start,\addrend,\addr,W) \land \\
-    &\quad  n' \leq n \\
-    &\qquad \Rightarrow n' \in \entryCond{}(\stdvr)(\start,\addrend,\addr,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Entry condition monotone in world]
-  \label{lem:en-mono-world}
-  \begin{align*}
-    &\forall \stdvr,W, W',\start, \addrend, \addr, \start', \addrend', \addr' \ldotp \\
-    &\quad  (\start',\addrend',\addr',W') \future (\start,\addrend,\addr,W) \\
-    &\qquad \Rightarrow \entryCond{}(\stdvr)(\start',\addrend',\perm',W') \supseteq \entryCond{}(\stdvr)(\start,\addrend,\perm,W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Entry condition non-expansive in worlds]
-\label{lem:en-ne-world}
-  \begin{align*}
-    &\forall \stdvr,n, W, W',\start,\addrend,\addr, \start', \addrend', \addr' \ldotp \\
-    &\quad (\start,\addrend,\perm,W) \nequal (\start',\addrend',\perm',W') \Rightarrow \\
-    &\qquad \Rightarrow \execCond{}(\stdvr)(\start,\addrend,\addr,W) \nequal \execCond{}(\stdvr)(\start',\addrend',\addr',W')
-  \end{align*}
-\end{lemma}
-Finally, we need to show that all the conditions are non-expansive, but we later want to use Banach's fixed-point theorem to define the value relation.
-For this we will need that the above conditions are contractive, and if they are contractive, then they are also non-expansive, so we show that each of the conditions are contractive:
-\begin{lemma}[Read condition contractive]
-  \label{lem:rc-con}
-  \begin{align*}
-    &\forall \stdvr, \stdvr', n \ldotp \\
-    &\quad \stdvr \nequal \stdvr' \Rightarrow \readCond{}(\stdvr) \nequal[n+1] \readCond{}(\stdvr')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Write condition contractive]
-  \label{lem:wc-con}
-  \begin{align*}
-    &\forall \stdvr, \stdvr', n \ldotp \\
-    &\quad \stdvr \nequal \stdvr' \Rightarrow \writeCond{}(\stdvr) \nequal[n+1] \writeCond{}(\stdvr')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Execute condition contractive]
-  \label{lem:ec-con}
-  \begin{align*}
-    &\forall \stdvr, \stdvr', n \ldotp \\
-    &\quad \stdvr \nequal \stdvr' \Rightarrow \execCond{}(\stdvr) \nequal[n+1] \execCond{}(\stdvr')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Entry condition contractive]
-  \label{lem:en-con}
-  \begin{align*}
-    &\forall \stdvr, \stdvr', n \ldotp \\
-    &\quad \stdvr \nequal \stdvr' \Rightarrow \entryCond{}(\stdvr) \nequal[n+1] \entryCond{}(\stdvr')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Write condition implies read condition]
-\label{lem:wc-implies-rc}
-  \begin{align*}
-    &\forall n, W, \start, \addrend \ldotp\\ 
-    &\quad \writeCond{n,W,\start,\addrend} \Rightarrow \readCond{n,W,\start,\addrend}
-  \end{align*}
-\end{lemma}
-
-\subsubsection{Value Relation}
-\label{subsubsec:value-relation}
-The value relation, is defined as follows:
-\begin{align*}
-  \stdvr & :  (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
-  \stdvr &\defeq \lambda \; V \ldotp \lambda \; W \ldotp 
-              \begin{aligned}[t]
-  & \{ \npair{i} \mid i \in \ints \} 
-    \union \\
-         & \{ \npair{\stdcap[\noperm] }  \} 
-           \union \\
-         & \{ \npair{\stdcap[\readonly] } \mid n \in \readCond{}(V)(\start,\addrend,W) \} \union \\
-         & \{ \npair{\stdcap[\readwrite] } \mid n \in \writeCond{}(V)(\start,\addrend,W) \} \union \\
-         & \{ \npair{\stdcap[\exec]} \mid \\
-         & \quad n \in \readCond{}(V)(\start,\addrend,W) \land \\
-         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \}
-           \union \\
-           % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
-         & \{ \npair{\stdcap[\entry]} \mid n \in \entryCond{}(V)(\start,\addrend,\addr,W) \} \union \\
-         & \{ \npair{\stdcap[\rwx]} \mid \\
-         & \quad n \in \writeCond{}(V)(\start,\addrend,W) \land \\
-         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \land \\
-         & \quad n \in \execCond{}(V)(\start,\addrend,\rwx,W) \}
-       \end{aligned}
-\end{align*}
-
-\begin{lemma}[Value relation uniformity]
-\label{lem:stdvr-dc}
-  \begin{align*}
-    & \forall V, W, n, n' \ldotp \\
-    & \quad n' \leq n \land n \in \stdvr(V)(W) \\
-    & \qquad \Rightarrow n' \in \stdvr(V)(W)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Value relation non-expansive in $V$]
-\label{lem:stdvr-ne-V}
-  \begin{align*}
-    & \forall V, V', n \ldotp \\
-    & \quad V \nequal V' \Rightarrow \stdvr(V) \nequal \stdvr(V')
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}[Value relation monotone in worlds]
-\label{lem:stdvr-mono-world}
-  \begin{align*}
-    & \forall V, W, W' \ldotp \\
-    & W' \future W \Rightarrow \stdvr(V)(W') \supseteq \stdvr(V)(W)
-  \end{align*} 
-\end{lemma}
-
-\begin{lemma}[Value relation non-expansive in worlds]
-\label{lem:stdvr-ne-worlds}
-  \begin{align*}
-    & \forall V, n, W, W' \ldotp \\
-    & W \nequal W' \Rightarrow \stdvr(V)(W) \nequal \stdvr(V)(W')
-  \end{align*} 
-\end{lemma}
-
-\begin{lemma}[Value relation contractive]
-\label{lem:stdvr-con}
-  \begin{align*}
-    & \forall V, V', n \ldotp \\
-    & \quad V \nequal V' \Rightarrow \stdvr(V) \nequal[n+1] \stdvr(V')
-  \end{align*}
-\end{lemma}
-
-Seeing as Lemma~\ref{lem:stdvr-con} gives us that $\stdvr$ is contractive, the Banach fixed-point theorem gives us a fixed-point for $\stdvr$.
-We will use this fixed point in the next section to define the logical relation we use in practice.
-
-\subsection{Logical Relation Definitions}
-\label{subsec:logical-relation-definitions}
-\lau{This section introduces the definitions, we started working with, so it is here so we don't have to rewrite everything.}
-Here we state the definitions, we will use for practical purposes.
-Assume $V$ is the fixed point given by Banach's fixed point theorem as described in Section~\ref{subsubsec:value-relation}.
-In general the definitions here will be the same as in Section~\ref{subsec:logical-relation-setup}, but but applied to $V$ when applicable.
-As they are essentially the same, we will reuse the names.
-
-\begin{align*}
-  \stdrr & \defeq \stdrr \; V \\
-  \stdkr & \defeq \stdkr \; V \\
-  \stder & \defeq \stder \; V \\
-  \stdvr & \defeq \stdvr \; V 
-\end{align*}
-
-\begin{align*}
-  \iota_{\start,\addrend} & \defeq \iota_{\start,\addrend} \; V \\
-  H_{\var{std}}          & \defeq H_{\var{std}} \; V
-\end{align*}
-
-For natural number $n$, addresses $\start$, $\addrend$, and $\addr$, permission $\perm$, and world $W$, we define the capability conditions as follows
-\begin{align*}
-  \readCond{n,W,\start,\addrend}        & \text{ iff } n \in \readCond{}(V)(\start,\addrend,W)\\
-  \writeCond{n,W,\start,\addrend}       & \text{ iff } n \in \writeCond{}(V)(\start,\addrend,W)\\
-  \execCond{n,W,\start,\addrend,\perm}  & \text{ iff } n \in \execCond{}(V)(\start,\addrend,p,W)\\
-  \entryCond{n,W,\start,\addrend,\addr} & \text{ iff } n \in \entryCond{}(V)(\start,\addrend,\addr,W)
-\end{align*}
-The well-formedness lemmas of course still hold for this new presentation.
-
-\subsection{Further Standard Regions}
-\label{subsec:further-standard-regions}
-We have already defined the standard region $\iota_{\start,\addrend}$ which ensures that all the words in the region are in the value relation.
-See Section~\ref{subsubsec:standard-region}.
-We here define another useful region, namely the one that requires all the words in the region to be the constant 0.
-\[
-  \iota_{\start,\addrend}^0 \; : \; \Regions
-\]
-\begin{align*}
-  \iota_{\start,\addrend}^0 & \defeq ((\start,\addrend),=,H_{\var{std}}^0) \\
-  H_{\var{std}}^0\; (\start,\addrend)\; W     & \defeq\left\{ \npair{\heapseg} \middle| 
-                                                     \begin{aligned}
-                                                       & \dom(\heapseg) = [\start,\addrend] \land \\
-                                                       & \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  
-                                                     \end{aligned} \right\}
-\end{align*}
-
-\begin{lemma}[$\iota_{\start,\addrend}^0$ is well-defined]\label{lem:std-z-region-wd}\forcenewline
-\begin{itemize}
-  \item $H_{\var{std}}^0$ is non-expansive.
-  \item For all addresses $\start$
-    and $\addrend$, $ H_{\var{std}}^0 \; (\start,\addrend)$ is monotone and non-expansive.
-  \item  $=$ is a reflexive and transitive relation.
-\end{itemize}
-\end{lemma}
-
-\subsection{Fundamental Theorem of Logical Relations}
-\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
-  For any $n \in \nats$, $W \in \Worlds$, $p\in \Perms$, and addresses $\var{base}$, $\var{end}$, and $a$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, a)} \in \stder(W)$.
-\end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
-\lau{With the current definition of the expression relation, an entry pointer will never be in the expression relation, but I guess we do not want that restriction.
-Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
-\dominique{I don't understand the above remark: why should an entry pointer be
-  in the expression relation, since it shouldn't be able to reach the pc?}
-\lau{In the ticket dispenser lemma, we need to make some assumptions about the adversary to be able to use the FTLR.
-What should these assumptions be?
-So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition from this.}
-\dominique{Something I find interesting here, is that the FTLR can be interpreted to say that x permissions have no real security value: if you have an arbitrary r or rw capability that is valid, then the FTLR says that it remains valid if we change it to rx or rwx (respectively).}
-\begin{proof}
-  The proof is by induction on $n$.
-
-If $n=0$, then it all boils down to an assumption where $(\reg, \hs \uplus \heap_f) \step[0] (\halted, \heap')$ for some $\reg$, $\hs$, $\heap_f$, and $\heap'$ which is not possible in our operational semantics.
-
-If $n > 0$, then we need to consider a lot of cases, but first we make a lot of preliminary assumptions:\\
-Let the world $W$, permission $p$, and addresses $b$, $e$, and $a$ be given. Assume
-\begin{align}
-  & p = \rwx \text{ and } \writeCond{n,W,b,e} \text{ or } \label{eq:ftlr:rwx-assump}\\
-  & p = \readonly \text{ and } \readCond{n,W,b,e} \label{eq:ftlr:rx-assump}
-\end{align}
-We need to show 
-\[
-  \npair{(p,b,e,a)} \in \stder(W)
-\]
-Let $n'$, $\reg$, $c$, $\hs$ be given where
-\begin{align}
-  & n' \leq n \label{eq:ftlr:n-leq} \\
-  & \npair[n']{\reg} \in \stdrr(W) \label{eq:ftlr:reg-in-stdrr}\\
-  & \npair[n']{c} \in \stdkr(W) \label{eq:ftlr:cont-in-stdkr}\\
-  & \heapSat[\hs]{n'}{W} \label{eq:ftlr:hs-heapsat}
-\end{align}
-For $\pc = (p,b,e,a)$ we need to show
-\[
-\npair[n']{(\reg[r_0 \mapsto c, \pcreg \mapsto \pc],\hs)} \in \observations(W)
-\]
-For convenience, we will define $\reg' = \reg[r_0 \mapsto c, \pcreg \mapsto \pc]$. To show the above, let $h_f$, $h'$, and $i$ be given where $i \leq n'$ and assume
-\begin{equation}
-  \label{eq:ftlr:conf-steps}
-  \Phi \step[i] (\halted, h')
-\end{equation}
-for $\Phi = (\reg',\hs \uplus h_f)$ then we need to show
-\begin{align}
-  & \exists W' \future W \ldotp \exists \hs' \ldotp \nonumber \\
-  & \quad h' =  \hs' \uplus h_f \text{ and } \label{eq:ftlr:final-heap-cond} \\
-  & \quad \heapSat[\hs']{n'-i}{W'} \label{eq:ftlr:heap-sat-cond}
-\end{align}
-
-If $i=0$, then we are trivially done as execution~\ref{eq:ftlr:conf-steps} is not possible.
-If $i>0$, then execution~\ref{eq:ftlr:conf-steps} takes at least one step, consider the first step of this execution.
-By the operational semantics, the first step is:
-\begin{equation*}
-  \Phi \step \sem{\decode(h(a))}(\Phi)
-\end{equation*}
-
-where $h = \hs \uplus h_f$. We may further assume
-\begin{align}
-  &\reg'(\pcreg) = (\perm,\start,\addrend,\addr) \nonumber \\
-  &\start \leq a \leq  \addrend \nonumber \\
-  &\perm \in \{\exec,\rwx\} \nonumber
-\end{align}
-For convenience define $\instr = \decode(h(a))$. We proceed by case on $\instr$.
-\newcommand{\case}[1]{{\bf Case } $\instr=#1$:\\}
-
-\case{\fail} This case is trivially true as it is not possible to step to a halting state from a failed step in any number of steps.
-By assumption~\ref{eq:ftlr:conf-steps}, the execution ends in a halting state.
-
-\case{\halt} 
-Here we have:
-\[
-  \Phi \step (\halted, (\reg',h))
-\]
-Pick $\hs' = \hs$ and $W' = W$.
-$h' = \hs \uplus h_f$, so condition~\ref{eq:ftlr:final-heap-cond} is satisfied.
-Using $\heapSat[\hs]{n'}{W}$ and the uniformity of heap satisfaction, we get $\heapSat[\hs]{n'-i}{W'}$ as $n' > n'-i$.
-
-\case{\jmp{r}}
-Here the first execution step is:
-\[
-  \Phi \step (\reg[r_0 \mapsto c, \pcreg \mapsto \updatePcPerm{\reg'(r)}],h)
-\]
-We now consider four cases 
-\begin{description}
-\item[Case $\reg'(r) = (\rwx,b',e',a')$:] In this case, we want to appeal to the induction hypothesis.
-To do so, we need to argue that we have $\writeCond{n'-1,b',e',a'}$.
-If $r=\pcreg$, then we have this by assumption \ref{eq:ftlr:rwx-assump} and uniformity.
-If $r \neq \pcreg$, then if $r= r_0$, then by assumption \ref{eq:ftlr:cont-in-stdkr}, we have $\npair[n']{c} \in \stdkr(W)$ which entails $\npair[n']{c} \in \stdvr(W)$ which together with uniformity gives the $\writeCond{}$ due to the permission.
-Otherwise, the same follows from assumption \ref{eq:ftlr:reg-in-stdrr} and uniformity.
-
-By IH we get 
-\begin{equation}
-  \npair[n'-1]{(\rwx,b',e',a')} \in \stder(W) \nonumber
-\end{equation}
-using $n'-1$, $\reg$, $c$, and $\hs$ and assumptions \ref{eq:ftlr:n-leq} - \ref{eq:ftlr:hs-heapsat} and the appropriate uniformity lemmas, we can use the above to get
-\begin{equation}
-  \npair[n'-1]{reg[r_0 \mapsto c, \pcreg \mapsto (\rwx,b',e',a')]} \in \observations(W) \nonumber
-\end{equation}
-This is the register file of the configuration after the first step of execution \ref{eq:ftlr:conf-steps}, so we know:
-\begin{equation}
-  (reg[r_0 \mapsto c, \pcreg \mapsto (\rwx,b',e',a')],h) \step[i-1] (\halted, h') \nonumber
-\end{equation}
-using this and that the register file is in the observation relation, we get a $W' \future W$ and $\hs'$ such that 
-\begin{align*}
-  &h' = \hs' \uplus h_f \\
-  &\heapSat[\hs']{n'-1-(i-1)}{W'}
-\end{align*}
-$W'$ and $\hs'$ satisfy exactly the properties, we need, so we are done.
-
-\item[Case $\reg'(r) = (\exec,b',e',a')$:] This case goes just like the one above with the exception that we need $\readCond{n'-1,b',e',a'}$ which we get in a similar fashion.
-
-\item[Case $\reg'(r) = (\entry,b',e',a')$:] In this case, we do not appeal to the induction hypothesis, but rather to the $\entryCond{}$.
-If $r=\pcreg$, then we have a contradiction with either \ref{eq:ftlr:rwx-assump} or \ref{eq:ftlr:rx-assump}, so it is safe to assume $r \neq \pcreg$.
-If $r=r_0$, then we rely on assumption~\ref{eq:ftlr:cont-in-stdkr} to get $\npair[n']{\entry,b',e',a'} \in \stdvr(W)$.
-In the case where $r \neq r_0$, we rely on assumption~\ref{eq:ftlr:reg-in-stdrr} to get $\npair[n']{\entry,b',e',a'} \in \stdvr(W)$, so we will have $\entryCond{n',W,b',e',a'}$.
-Using the $\entryCond{}$, we get $\npair[n'-1]{(\exec,b',e',a')}\in \stder(W)$.
-From here on, the argument follows like the cases, we have seen.
-One notable thing is that in this case is that $\updatePcPerm{\reg[r_0 \mapsto c, \pcreg \mapsto \pc](r)}$ updates the permission to an execute permission $\exec$ which makes everything align up nicely as in the previous cases.
-
-\item[Otherwise] in this case, we consider the next execution step which goes to $\failed$ as the $\pcreg$ is not a well-formed execution.
-We have, however, assumed that a halting configuration is reached eventually, so a contradiction is reached.
-
-\end{description}
-
-\case{\jnz{}{}}
-\lau{TODO, I have not checked the details as I expect it to be like the $\jmp{}$ case.}
-
-\case{\load{r_1}{r_2}}
-Assume for some $p'$, $b'$, $e'$, $a'$, and $w$ that
-\begin{align}
-  &\reg'(r_1) = (p',b',e',a') \nonumber\\
-  &\readAllowed{p'} \label{eq:ftlr:load-read-allowed} \\
-  &\withinBounds{(p',b',e',a')} \nonumber \\
-  &w = h(a) \nonumber 
-\end{align}
-if the above where not the case, then by the semantics, we would step to a fail configuration which by assumption is not possible. We consider two cases for $r_1$.
-\begin{description}
-\item[$r_1 = \pcreg$] The first step of the execution is
-\[
-  \Phi \step \stdUpdatePc{\updateReg{\pcreg}{w}}
-\]
-If $w$ is not a capability, then
-\[
-\stdUpdatePc{\updateReg{\pcreg}{w}} = \failed
-\]
-and we are done. Assume $w = (p',b',e',a')$, then
-\[
-\stdUpdatePc{\updateReg{\pcreg}{w}} = \updateReg{\pcreg}{(p',b',e',a'+1)}
-\]
-Using execution~\ref{eq:ftlr:conf-steps}, we can conclude that $p' \in \{\rwx,\exec\}$ as the above configuration otherwise would step to $\failed$. From \ref{eq:ftlr:load-read-allowed}
-
-\item[$r_1 \neq \pcreg$]
-\end{description}
-
-\case{\store{}{}}
-
-\case{\move{}{}}
-
-\case{\lea{}{}}
-
-\case{\restrict{}{}{}}
-
-\case{\plus{}{}{}}
-
-\end{proof}
-
 \section{Other examples and applications}
@@
 
-\begin{appendices}
-
-\section{Proofs of Lemmas in Section~\ref{sec:logical-relation}}
-\subsection{Definitions and Proofs related to Section~\ref{subsec:recursive-dom-eq}}
-
-\begin{definition}[Ordering of Cauchy chains]
-  For Cauchy chains $\{a_n\}_n$ and $\{b_n\}_n$,
-  \begin{align*}
-    \{a_n\}_n \geq \{b_n\}_n & & \text{iff} & & \forall n \ldotp a_n \geq b_n
-  \end{align*}
-\end{definition}
-
-\begin{definition}[Limit of Cauchy chains]
-  For a Cauchy chain $\{a_n\}_n$ in the \cofe{}\ $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$, we say that $a$ is a limit  (denoted $\lim\{a_n\}_n$) of this chain iff it satisfies
-  \begin{align*}
-    & \forall n \ldotp \exists N \ldotp \forall m \geq N \ldotp a \nequal a_m
-  \end{align*}
-\end{definition}
-
-\begin{definition}[Set \cofe{}\ Construction]
-  Given a set $X$, then $\left(X, \left(\nequal\right)_{n=0}^{\infty} \right)$ is a \cofe{}\ where all the equivalences are equality.
-\end{definition}
-
-\begin{definition}[Function \cofe{}\ Construction]
-  Given \cofes{} $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the \cofe{}\ $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) \fun \left(X, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ as
-\[
-\left(\{f : X \fun Y \mid f \text{ non-expansive}\}, \left(\nequal\right)_{n=0}^{\infty} \right)
-\]
-where the equivalence is defined as
-\begin{align*}
-  f \nequal g & & \text{ iff } & & \forall x \in X \ldotp f(x) \nequal_Y g(x)
-\end{align*}
-
-\end{definition}
-
-\begin{proof}[Proof of Lemma~\ref{lem:reg-def-wd}]
-  The \cofe{}\ is a standard construction for tuples, so we won't show that it in fact is a \cofe{}\ here.
-That $\future$ is reflexive and transitive follows trivially from $\Phi$ being reflexive and transitive.
-It remains to show that $\future$ preserves limits.
-
-  Assume $\{a_n\}_n \geq \{b_n\}_n$.From this it follows that for all $m \in \nats$:
-  \begin{align*}
-    \Phi_{b_m} & = \Phi_{a_m} \\
-    H_{b_m} & = H_{a_m} \\
-    (s_{a_m},s_{b_m}) &\in \Phi_{b_m}
-  \end{align*}
-  Say $\lim{a_n}_n = (s_a,\Phi_a,H_a)$ and $\lim{b_n}_n = (s_b,\Phi_b, H_b)$. We need to show
-  \begin{align*}
-    \Phi_{b} & = \Phi_{a} \\
-    H_{b} & = H_{a} \\
-    (s_{a},s_{b}) &\in \Phi_{b}
-  \end{align*}
-  To show $H_{b} = H_{a}$ by one of the \cofe{}\ properties SFTS $\forall n \ldotp H_{b} \nequal H_{a}$. Given $n$ use that $a$ is a limit to get $N$ s.t.
-\[
-  \forall m \geq N \ldotp H_{a_m} \nequal H_a
-\]
-Likewise for $b$ we get a similar $M$. Take some $m \geq \max\{M,N\}$, using $H_{b_m} = H_{a_m}$, we get:
-\[
-  H_a = H_{a_m} = H_{b_m} = H_b
-\]
-
-For $\Phi_{b} = \Phi_{a}$ we do something similar except, we just pick $n=1$ as the $n$-equality in this \cofe{} is equality. 
-Same approach is used for $(s_{a},s_{b}) \in \Phi_{b}$
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:world-finpar}]
-\[
-  K(f)(\phi) = \lambda n \ldotp f(\phi(n))
-\]
-The above is defined exactly on the $n\in \nats$ where $\phi$ is defined. $\phi$ is a finite partial function, so the above is as well.
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:HSP-mon}]
-  Given $g$, $H$, $\var{st}$, $w$ and $w'$. Assume $w' \future w$. Show
-  \begin{gather}
-    (G(g))(\phi)(\var{st})(w') \future (G(g))(\phi)(\var{st})(w) \nonumber\\ 
-    \Uparrow \label{lem:HSP-mon:b1}\\
-    H(\var{st})(g(w')) \future H(\var{st})(g(w)) \nonumber \\
-    \Uparrow \label{lem:HSP-mon:b2}\\
-    g(w') \future g(w) \nonumber\\
-    \Uparrow \label{lem:HSP-mon:b3}\\
-    w' \future w \nonumber
-  \end{gather}
-(\ref{lem:HSP-mon:b1}) expands the definition of $G$.
-(\ref{lem:HSP-mon:b2}) is due to $H(\var{st})$ being monotone.
-(\ref{lem:HSP-mon:b3}) is due to $g$ being a morphism in $\CatC^{\var{op}}$ and they are monotone function.
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:HSP-ne}]
-  Given $n$, $g$, $H$, $\var{st}$, $w$ and $w'$. Assume $w' \nequal w$. Show
-  \begin{gather}
-    (G(g))(\phi)(\var{st})(w') \nequal (G(g))(\phi)(\var{st})(w) \nonumber\\ 
-    \Uparrow \nonumber \\
-    H(\var{st})(g(w')) \nequal H(\var{st})(g(w)) \nonumber \\
-  \end{gather}
-  We use that $H=H$ implies $H \nequal H$ which means
-  \[
-    \forall \var{st}, \var{st'} \ldotp \var{st} \nequal \var{st'} \Rightarrow H(\var{st}) \nequal H(\var{st'})
-  \]
-  This in turn implies
-  \[
-    \forall y,y' \ldotp \ldotp y \nequal y' \Rightarrow H(\var{st})(y) \nequal H(\var{st})(y')
-  \]
-  Using our assumption, about the worlds, we are done.
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:K-func}]
-  $K(f) : K(X) \rightarrow K(Y)$ is \emph{monotone} for $f : X \monnefun Y$:
-  Assume $\phi \future \phi'$. Show
-  \[
-    (K(f))(\phi) \future (K(f))(\phi') 
-  \]
-  By def on ordering for functions SFTS
-  \[
-    \forall n \ldotp (K(f))(\phi)(n) \future (K(f))(\phi')(n)
-  \]
-  let $n$ be given show
-  \[
-    f(\phi'(n)) \future f(\phi(n))
-  \]
-  $f$ mon, so SFTS
-  \[
-    \phi'(n) \future \phi(n)
-  \]
-  which follows from the assumption $\phi' \future \phi$.\\\\
-  $K(f) : K(X) \rightarrow K(Y)$ is \emph{non-expansive} for $f : X \monnefun Y$:
-  Assume $\phi \nequal \phi'$. Show
-  \[
-    (K(f))(\phi) \nequal (K(f))(\phi') 
-  \]
-  Let $k$ be given, SFTS
-  \[
-    (K(f))(\phi)(k) \nequal (K(f))(\phi')(k)
-  \]
-  which unfolds to
-  \[
-    f(\phi(k)) \nequal f(\phi'(k))
-  \]
-  as $f$ is non-expansive SFTS
-  \[
-    \phi(k) \nequal \phi'(k)
-  \]
-  which follows from assumption.\\\\
-  $K$ preserves composition, show
-  \[
-    K(f \circ g) = K(f) \circ K(g)
-  \]
-  Given $\phi$ and $n$ show
-  \begin{align*}
-    (K(f) \circ K(g))(\phi)(n) & = K(f)(K(g)(\phi))(n) \\
-                               & = f(K(g)(\phi)(n)) \\
-                               & = f(g(\phi(n))) \\
-                               & = (f \circ g)(\phi(n)) \\
-                               & = K(f \circ g)(\phi)(n)
-  \end{align*}\\\\
-  $K$ preserves identity, show
-  \[
-    K(\id) = \id
-  \]
-  Let $\phi$ and $n$ be given.
-  \begin{align*}
-    K(\id) & = \lambda \phi \ldotp \lambda n \ldotp \id(\phi(n)) \\
-           & = \lambda \phi \ldotp \lambda n \ldotp \phi(n) \\
-           & = \lambda \phi \ldotp \phi \\
-           & = \id
-  \end{align*}
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:R-func}]
-   $R(f) : R(X) \rightarrow R(Y)$ is \emph{monotone} for $f : X \nefun Y$. Given aforementioned $f$. Further let $(s_2,\Phi_2,H_2) \future (s_1,\Phi_1,H_1)$ be given.
-
-    \begin{gather*}
-      R(f)(s_2,\Phi_2,H_2) = (s_2,\Phi_2,H_2(f)) \future (s_1,\Phi_1,H_1(f)) = R(f)(s_1,\Phi_1,H_1) \\
-      \Uparrow\\
-      \Phi_2 = \Phi_1 \text{ and } H_1(f) = H_2(f) \text{ and } (s_1,s_2) \in \Phi_2
-    \end{gather*}
-    From our assumption, we have $H_2 = H_1$ from which it follows $H_1(f) = H_2(f)$. The remaining conditions also follow directly from the assumption.
-
-    $R(f) : R(X) \rightarrow R(Y)$ is \emph{non-expansive} for $f : X \nefun Y$. Let $(s_2,\Phi_2,H_2) \nequal (s_1,\Phi_1,H_1)$ be given
-\[
-  R(f)(s_2,\Phi_2,H_2) = (s_2,\Phi_2,H_2(f)) \nequal (s_1,\Phi_1,H_1(f)) = R(f)(s_1,\Phi_1,H_1)
-\]
-$H_1(f) \nequal H_2(f)$ follows from the definition of $n$-equality on functions as well as the fact that $f=f$ implies $f \nequal f$ for any $n$.
-
-  Composition distributes over $R$, $R(f \circ g) = R(f) \circ R(g)$.
-  \begin{align*}
-    R(f) \circ R(g)(s,\Phi,H) &= R(f)(s,\Phi,g(H)) \\
-                              &= (s,\Phi,f(g(H)))\\
-                              &= R(f \circ g)
-  \end{align*}
-
-  Identity distributes over $R$, $R(\id) = \id$.
-  \[
-    R(id) = \lambda (s,\Phi,H) \ldotp  (s,\Phi,\id(H)) = \lambda (s,\Phi,H) \ldotp  (s,\Phi,H) = \id
-  \]
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:G-func}]
-  $G(f) : G(Y) \rightarrow G(X)$ is \emph{non-expansive} for $f : X \monnefun Y$. For $H \nequal H'$ show
-  \[
-    G(f)(H) \nequal G(f)(H')
-  \]
-  which expands to
-  \[
-    \lambda \var{st} \ldotp \lambda w \ldotp H(\var{st})(f(w)) \nequal \lambda \var{st} \ldotp \lambda w \ldotp  H'(\var{st})(f(w))
-  \]
-  Given $\var{st} \nequal \var{st}'$ and $w \nequal n'$ SFTS
-  \[
-  \ldotp H(\var{st})(f(w)) \nequal H'(\var{st}')(f(w'))
-  \]
-  As $f$ is non-expansive, we get $f(w) \nequal f(w')$. As $H \nequal H'$ we get $H(\var{st}) \nequal H(\var{st'})$. Combining the two, we get the desired result.  
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:F-func}]
-Follows from Lemmas~\ref{lem:K-func},~\ref{lem:R-func}, and~\ref{lem:G-func}.
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:F-loc-ne}]
-Let $k$, $f$ and $g$ be given and assume $f \nequal[k] g$. Show
-\[
-F(f) \nequal[k] F(g).
-\]
-SFTS
-\[
-\forall \phi \ldotp (F(f))(\phi) \nequal[k] (F(g))(\phi)
-\]
-Let $\phi$ be given. We need to show $\dom((F(f))(\phi)) = \dom((F(g))(\phi))$. Both sides are defined exactly when $\phi$, so they have the same domain. We also need to show
-\[
-\forall n \in \dom((F(f))(\phi))\ldotp (F(f))(\phi)(n) \nequal[k] (F(g))(\phi)(n)
-\]
-Let $n \in  \dom((F(f))(\phi))$ be given. Define $H = \pi_3 (\phi(n))$. The above unfolds to two triples, so SFTS each component is $k$-equal. That is
-\begin{align*}
-  \pi_1(\phi(n)) \nequal[k] \pi_1(\phi(n))\\
-  \pi_2(\phi(n)) \nequal[k] \pi_2(\phi(n))\\
-  \lambda \var{st}'\ldotp (H(\var{st}') \circ f) \nequal[k]   \lambda \var{st}'\ldotp (H(\var{st}') \circ g)
-\end{align*}
-The two first are trivially satisfied. For the third, we need to show that for all $\var{st}$ if we apply $\var{st}$ to each side, then they are still $k$-equal. That is given $\var{st}$ show
-\[
-  (H(\var{st}) \circ f) \nequal[k]   \ldotp (H(\var{st}) \circ g)
-\]
-SFTS
-\[
-\forall w \ldotp (H(\var{st}) \circ f)(w) \nequal[k]   \ldotp (H(\var{st}) \circ g)(w)
-\]
-Let $w$ be given. From $f \nequal[k] g$, we get $f(w) \nequal[k] g(w)$. From this and $H(st)$ being non-expansive, we get
-\[
-  H(\var{st})(f(w)) \nequal[k]   H(\var{st})(g(w))
-\]
-which is what we needed to show.
-
-\end{proof}
-\subsection{Proofs of Lemmas in Section~\ref{subsec:worlds}}
-%TODO All proofs written by hand, typeset them.
-\begin{proof}[Proof of lemma~\ref{lem:nequal-and-future-world}]
-  Assume $W_1 \nequal W_2$ and $W_1' \future W_2'$. Define $W_2'$ to have $\dom(W_2')=\dom(W_1')$ where
-\[
-  W_2' \; r = \left\{
-  \begin{aligned}
-    & (s_1',\phi_2,H_2) & &\text{for } j \in \dom(W_2) \text{ and } W_2 \; j = (s_2,\phi_2,H_2) \\
-    &                  & &\quad \text{ and } W_1' \; j = (s_1',\phi',H_1') \\
-    & W_1' \; r        & &\text{otherwise}
-  \end{aligned} 
-  \right.
-\]
-
-First we show $W_1' \nequal W_2'$.
-The domains are defined to be equal, so that is satisfied.
-Let $r \in \dom(W_1')$ and show $W_1' \; r \nequal W_2' \; r$.
-First consider the case where $r \in \dom(W_2)$.
-Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$.
-We need to show
-\begin{align*}
-  s_1' & \nequal s_1' \\
-  \phi_1' & \nequal \phi_2 \\
-  H_1' & \nequal H_2
-\end{align*}
-The first one is trivial.
-For the second one we use that we know $\phi_1 \nequal \phi_2$ and $\phi_1 = \phi_1'$ respectively from the $n$-equality assumption and future world assumption.
-Similarly, the second one follows from $H_1 \nequal H_2$ and $H_1 = H_1'$ - again respectively from the $n$-equality assumption and the future world assumption.
-For the case where $r \not\in \dom(W_2)$, we need to show $W_1' \; r \nequal W_1' \; r$ which follows trivially from reflexivity.
-
-Next we show $W_2' \future W_2$.
-We know that $\dom(W_2') = \dom(W_1')$, $\dom(W_1') \supseteq \dom(W_1)$ (from future world assumption), and $\dom(W_1) = \dom(W_2)$ (from $n$-equality assumption).
-All this can easily be put together to get $\dom(W_2') \supseteq \dom(W_2)$.
-Now let $r\in \dom(W_2)$ and assume Assume $W_1' \; r = (s_1',\phi_1',H_1')$, $W_1 = (s_1,\phi_1,H_1)$, and $W_2 = (s_2,\phi_2,H_2)$, then $W_2' \; r = (s_1',\phi_2,H_2)$, then we need to show the following:
-\begin{align*}
-  &(s_2,s_1') \in \phi_2 \\
-  &\phi_2 = \phi_2 \\
-  &H_2 = H_2
-\end{align*}
-Here the two last are trivial.
-The first follows from the assumptions $s_1 = s_2$, $\phi_1 = \phi_2$, and $(s_1,s_1') \in \phi_1$, where the first two are from the $n$-equality assumption (notice that $n$-equality between \cofes{} over sets is just equality) and the last one follows from the future world assumption.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-dc}]
-  Follows directly from definition: Let $\stdvr$, $n$, $n'$, $\start$, and  $\addrend$ be given.
-Assume $n \in \readCond{}(\stdvr)(\start,\addrend,W)$ and $n' \leq n$.
-If $n=n'$, then the result follows trivially, as it is exactly our assumption.
-If $n' < n$, then by assumption we have a region name $r$, interval $[\start',\addrend'] \supseteq [\start,\addrend]$, and the assumption $W(r) \nsubeq[n-1] \iota_{\start',\addrend'}$  given by the read condition assumption.
-Picking $r$ and $[\start',\addrend']$, we need to show  $W(r) \nsubeq[n'-1] \iota_{\start',\addrend'}$ which follows from downwards closure of $\nsubeq[n-1]$.
-\end{proof}
-
-\subsection{Proofs of Lemmas in Section~\ref{subsec:logical-relation-setup}}
-%All the lemmas are in subsections.
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:observation-relation}}
-%TODO All proofs written by hand (but one), typeset them.
-\begin{proof}[Proof of lemma~\ref{lem:obs-dc}]
-  Let $n' < n$, $W$, $\reg$, and $\hs$ be given.
-Assume $\npair{(\reg,\hs)} \in \observations(W)$.
-Let $\heap_f$, $heap'$, $i \leq n'$ be given and assume $(\reg, \hs \uplus \heap_f) \step[i] (\halted, \heap')$.
-By assumption, we have a $W' \future W$ and $\hs'$ such that
-  \begin{align}
-&\heap' = \hs' \uplus \heap_f \label{lem:eq:stdvr1}\\
-&\heapSat[\hs']{n-i}{W'}\label{lem:eq:stdvr2}
-  \end{align}. 
-Using $W'$ and $\hs'$ as existential witnesses, we already have Equation~\ref{lem:eq:stdvr1} as the first necessary condition and from the above heap satisfaction along with heap satisfaction being uniform in $n$, we get $\heapSat[\hs']{n'-i}{W'}$.
-These are the two conditions necessary to get $\npair[n']{(\reg,\hs)} \in \observations(W)$.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:obs-ne-worlds}]
-  %TODO
-\end{proof}
-
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:register-file-relation}}
-%TODO All proof but one proof written by hand, typeset them.
-%TODO missing proof register relation ne in worlds.
-\begin{proof}[Proof of lemma~\ref{lem:reg-dc}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:reg-mono-worlds}]
- %TODO,  This follows by monotonicity of $\stdvr$ in the worlds.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:reg-ne-vr}]
-  %TODO
-\end{proof}
-
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:continuation-relation}}
-%TODO All proof but one proof written by hand, typeset them.
-%TODO missing proof continuation relation ne in worlds.
-\begin{proof}[Proof of lemma~\ref{lem:stdkr-dc}]
-  % TODO, Follows by using the assumption and transitivity of inequality.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdkr-mono-worlds}]
-  % TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdkr-ne-vr}]
-  %TODO
-\end{proof}
-
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:expression-relation}}
-%TODO All proofs written by hand, typeset them.
-\begin{proof}[Proof of lemma~\ref{lem:stder-dc}]
-  %TODO,  Follows from the assumption.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stder-ne-worlds}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stder-ne-vr}]
-  %TODO
-\end{proof}
-
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:standard-region}}
-%TODO All proofs written by hand, typeset them.
-\begin{proof}[Proof of Lemma~\ref{lem:std-reg-mono-world}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:std-reg-ne-world}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:std-reg-ne-vr}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of Lemma~\ref{lem:std-reg-i-ne-vr}]
-  %TODO
-\end{proof}
-
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:capability-conditions}}
-%TODO All proofs written by hand, typeset them.
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-mono-world}]
-%TODO
-\end{proof}
-\begin{proof}[Proof of lemma~\ref{lem:rc-ne-world}]
-%TODO
-\end{proof}
-\begin{proof}[Proof of lemma~\ref{lem:wc-dc}]
-%TODO, Follows directly from the definition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:wc-mono-world}]
-%TODO, Similar to the proof of the monotonicity of the read condition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-dc}]
-%TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-mono-world}]
-%TODO, Follows directly from definition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-ne-world}]
-%TODO 
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-dc}]
-  Like the $\execCond{}$ uniformity proof.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-mono-world}]
-  Follows directly from definition.
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-ne-world}]
-  This proof goes like the proof of $\execCond{}$ being non-expansive in the second argument (lemma~\ref{lem:ec-ne-world}).
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:rc-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:wc-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:ec-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:en-con}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:wc-implies-rc}]
-  Follows directly from the definition.
-\end{proof}
-
-\subsubsection{Proofs of Lemmas in Section~\ref{subsubsec:value-relation}}
-%TODO All proofs written by hand, typeset them.
-\begin{proof}[Proof of lemma~\ref{lem:stdvr-dc}]
-  Follows from the uniformity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$.
-  %TODO show one case to illustrate the above, refer to the lemmas in question
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdvr-ne-V}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdvr-mono-world}]
-  Follows from monotonicity of $\readCond{}$, $\writeCond{}$, $\execCond{}$, and $\entryCond{}$ in the worlds.
-That is Lemma~\ref{lem:rc-mono-world},~\ref{lem:wc-mono-world},~\ref{lem:ec-mono-world}, and~\ref{lem:en-mono-world}.
-  %TODO show one case to illustrate the above
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdvr-ne-worlds}]
-  %TODO
-\end{proof}
-
-\begin{proof}[Proof of lemma~\ref{lem:stdvr-con}]
-  %TODO
-\end{proof}
-
-\subsection{Proofs of Lemmas in Section~\ref{subsec:further-standard-regions}}
-%TODO all the proofs need to be done.
-
-\begin{proof}[Proof of Lemma~\ref{lem:std-z-region-wd}]
-  %TODO
-\end{proof}
-\end{appendices}
-
 \end{document}

### 2016-11-18 local-capability-machine.tex
Added something about the stack and figures to illustrate how it works

@@
 Jumps to $\texttt{f}$.
-\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory. In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
-\item[\texttt{hstore x n}] Fetches the capability refered to with \texttt{x} and stores \texttt{n} to it.
+\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
+\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
 \item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them.
@@
 \end{description}
+
+\subsection{Stack}
+Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma.
+The stack is accessed through a local $\rwlx$-capability.
+Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
+\begin{description}
+\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
+\item[\texttt{pop r}] Pops the top word of the stack by loading it to regiser $r$, moves 0 to the address, and decrementing the stack pointer.
+\item[\texttt{scall f(r)}] This call assumes the presence of a stack!
+Push the code that restores the stackpointer, $\pcreg$ and $r_0$.
+Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer.
+Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$.
+Restrict the stackpointer to the ``unused part''.
+Push the arguments to the stack and jump to \texttt{f}.
+  \begin{description}
+  \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code.
+  \end{description}
+Figure \ref{fig:stack-before-call}-\ref{fig:stack-after-call} illustrates what the stack would look like at different stages of the scall.
+\end{description}
+
+\begin{figure}
+  \label{fig:stack-before-call}
+  \centering
+  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
+$r_{\var{stk}} \rightarrow$   & local state\\
+   & $\vdots$
+\end{tabular}
+\caption{Stack after everything stored}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
+$r_{\var{stk}} \rightarrow$     & Stack cap\\
+\cline{2-2}
+   & $r_0$-continuation\\
+\cline{2-2}
+   & o-$\pcreg$\\
+\cline{2-2}
+   & Restore code\\
+   & $\vdots$\\
+\cline{2-2}
+   & local state\\
+   & $\vdots$
+\end{tabular}
+\caption{Stack after everything stored}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
+$r_{\var{stk}} \rightarrow$      & arg1 \\
+\cline{2-2}
+   & $\vdots$\\
+\cline{2-2}
+   & argn \\
+\cline{2-2}
+   & Stack cap\\
+\cline{2-2}
+   & $r_0$-continuation\\
+\cline{2-2}
+   & o-$\pcreg$\\
+\cline{2-2}
+   & Restore code\\
+   & $\vdots$\\
+\cline{2-2}
+   & local state\\
+   & $\vdots$
+\end{tabular}
+\caption{Stack after arguments pushed to stack an ready to jump}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+   & ? \\
+\cline{2-2}
+   & Stack cap\\
+\cline{2-2}
+   & $r_0$-continuation\\
+\cline{2-2}
+   & o-$\pcreg$\\
+\cline{2-2}
+$\pcreg \rightarrow$   & Restore code\\
+   & $\vdots$\\
+\cline{2-2}
+   & local state\\
+   & $\vdots$
+\end{tabular}
+\caption{Stack just after return}
+\end{figure}
+
+\begin{figure}
+  \label{fig:stack-after-call}
+  \centering
+  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
+$r_{\var{stk}} \rightarrow$   & local state\\
+   & $\vdots$
+\end{tabular}
+\caption{Stack after everything has been restored}
+\end{figure}
+\subsection{Labels}
+\texttt{l:} defines a meta level label.
 
@@
 \subsubsection{Simple use of local capabilities}
-ML-like program:
+Old ML-like program:
 \begin{verbatim}
@@
 \begin{verbatim}
-f:  malloc l
+f1: malloc l
     hstore l 1
@@
     assert l 1
+    halt
+1f:
 \end{verbatim}

### 2016-11-18 local-capability-machine.tex
lemma

@@
 \usepackage{tikz}
-\usepackage{multirow}
+\usepackage{multirow,bigdelim}
 
@@
 \begin{description}
-\item[\texttt{call f(r)}] Moves the argument to register $\texttt{r}$ (if there are more arguments, then they are moved to the appropriate registers).
+\item[\texttt{call f(r)}] Moves the argument to a register (if there are more arguments, then they are moved to the appropriate registers).
 Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return).
-Erases all registers but the argument register(s), $\pcreg$, and the return register.
+Erases all registers but the argument register(s), $\pcreg$, the $\malloc$ register (if there is one), and the return register.
 Jumps to $\texttt{f}$.
@@
 \subsection{Labels}
-\texttt{l:} defines a meta level label.
+\texttt{l:} is a meta level label that can be used to refer to a specific address.
 
@@
 \subsubsection{Simple use of local capabilities}
-Old ML-like program:
+ML-like program we had before:
 \begin{verbatim}
@@
 1f:
+\end{verbatim}
+\begin{lemma}[Correctness lemma for \texttt{f1}]
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
+  \end{align*}
+if
+  \[
+    (\reg_0\update{\pcreg}{c_{\adv}}
+          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_f) \step[n] (\halted,h')
+  \]
+where 
+\begin{align*}
+  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
+  &\dom(\hs_\flag) = [\flag,\flag] \\
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]}
+\end{align*}
+and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
+then
+\[
+  h'(\flag) = 0
+\]  
+\end{lemma}
+
+Assembly program using the stack:
+\begin{verbatim}
+f2: push 1
+    scall adv()
+    pop r1
+    assert r1 1
+    halt
+2f:
 \end{verbatim}
@@
 \end{verbatim}
+\lau{ This example relies on \texttt{adv} actually passing on a proper $\malloc$ - is that a fair assumption?
+The example can be changed to not require this by assuming x and the function is already in the heap.
+This change does not remove the difficulties in the original program.
+}
 \begin{lemma}\forcenewline

### 2016-11-21 local-capability-machine.tex
Some remarks and questions...

@@
 In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
+\dominique{how is the return capability constructed: not on the stack, I presume, so this means mallocing a block of memory for the activation record?}
+\dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
+\dominique{what does ``fetch the capability ...'' mean?}
+\dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e.
+which do not need to be erased)?}
+\dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}
 \begin{description}
@@
 Push the code that restores the stackpointer, $\pcreg$ and $r_0$.
-Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer.
+Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?}.
 Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$.
@@
   \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code.
+\dominique{restores other registers too, right?}
   \end{description}
@@
   \end{align*}
-if
+if \dominique{$\hs_f$ \emph{and} $\hs_{f1}$?}
   \[
@@
   &\dom(\hs_\flag) = [\flag,\flag] \\
-  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]}
-\end{align*}
-and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
+\end{align*}
+and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ \dominique{do you mean f1?} including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
 then

### 2016-11-21 local-capability-machine.tex
Added some comments

@@
 \dominique{how is the return capability constructed: not on the stack, I presume, so this means mallocing a block of memory for the activation record?}
+\lau{See the new addition below - when we don't have a stack, then we have no need to ``save'' local capabilities which means that we can have the return code as part of the program.
+The only thing we have to do is to have the capability point to the return address and make the capability into an entry capability.
+}
 \dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
+\lau{ I think this is a conceptual question as we can make it work with either.
+We already trust malloc to give us a fresh piece of memory and not reuse it later on, so we already assume malloc to be somewhat trusted, so why not go all the way?
+}
 \dominique{what does ``fetch the capability ...'' mean?}
+\lau{ We don't know where the capability resides, but if it is in memory, then it will be loaded into a register. }
 \dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e.
@@
 Jumps to $\texttt{f}$.
+  \begin{description}
+  \item[return capability] The return capability is an enter capability into caller code.
+Upon return, the caller may assume that any return value is stored in $r_1$.
+The caller is responsible for clearing the remaining registers as part of its program.
+  \end{description}
 \item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
@@
    & $\vdots$
+\end{tabular}
+\caption{Stack after arguments pushed to stack an ready to jump}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{tabular}[!h]{r|>{\centering\arraybackslash}p{3cm}|l}
+   & &\\
+   & $\vdots$ &\\
+\cline{2-2}
+   & 0 &\\
+\cline{2-2}
+$r_{\var{stk}} \rightarrow$      & arg1 & \\
+\cline{2-2}
+   & $\vdots$ & \\
+\cline{2-2}
+   & argn & \\
+\cline{2-2}
+   & Stack cap & \\
+\cline{2-2}
+   & $r_0$-continuation &  \\
+\cline{2-2}
+   & o-$\pcreg$ & \\
+\cline{2-2}
+   & Restore code & \rdelim\}{2}{2cm}[variable length]\\
+   & $\smash{\vdots}$ & \\
+\cline{2-2}
+   & & \\
+\cline{2-2}
+   & local state & \\
+   & $\smash{\vdots}$ &
 \end{tabular}

### 2016-11-21 local-capability-machine.tex
Added malloc specification to new file. Changed call macro.

@@
 
+\section{Malloc specification}
+\newcommand{\hsfoot}{\hs_\var{footprint}}
+\newcommand{\hsframe}{\hs_\var{frame}}
+\newcommand{\size}{\var{size}}
+\newcommand{\rio}{r_{io}}
+\renewcommand{\adv}{\codelabel{adv}}
+\newcommand{\advb}{\var{adv_{base}}}
+\newcommand{\adve}{\var{adv_{end}}}
+\newcommand{\initb}{\var{init}_{base}}
+\newcommand{\inite}{\var{init}_{end}}
+\newcommand{\mrlen}{5cm}
+\newcommand{\retm}{\var{ret}_{\malloc}}
+\newcommand{\reta}{\var{ret}_{\adv}}
+\newcommand{\base}{\var{base}}
+\newcommand{\eend}{\var{end}}
+\newcommand{\bracket}[1]{\multirow{#1}{*}{\ensuremath{
+ \left . \vphantom{\begin{array}{l}
+ \ifthenelse{\equal{#1}{1}}{3\\}{
+    \ifthenelse{\equal{#1}{2}}{3\\3\\}{
+    \ifthenelse{\equal{#1}{3}}{3\\3\\3\\}{
+    \ifthenelse{\equal{#1}{4}}{3\\3\\3\\3\\}{
+    \ifthenelse{\equal{#1}{5}}{3\\3\\3\\3\\3\\}{
+    \ifthenelse{\equal{#1}{6}}{3\\3\\3\\3\\3\\3\\}{
+      3\\3\\3\\3\\3\\3\\3\\ %7
+  }}}}}}
+  \end{array}} \right \}}}
+}
+\newcommand{\annotate}[2]{\multirow{#1}{\mrlen}{\scriptsize #2}}
+
+\begin{specification}[Malloc v.2]
+  $c_\malloc$ satisfies the specification for malloc iff
+  \begin{align*}
+    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
+    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
+    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
+    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
+%    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\ % This
+    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
+    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
+    &\qquad \quad \memreg(r_1) = \size \land \\
+    &\qquad \quad \size \geq 0 \land \\
+    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
+    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
+    &\qquad \quad \Rightarrow \\
+    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
+    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
+    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
+    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot'\uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
+    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
+    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
+    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
+    &\qquad \qquad \qquad \size = e'-b' \\
+    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
+    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
+  \end{align*}
+
+  Additionally, we require that for every $W' \future [i\mapsto \iota_{\malloc,0}]$, $\heapSat[\hs]{n}{W'}$ implies that $\dom(\hs)$ contains $[n',+\infty)$ for some $n'$.
+\end{specification}
+
 \section{Macros}
@@
 The only thing we have to do is to have the capability point to the return address and make the capability into an entry capability.
+}
+\lau{On further thought, conveniently the examples did not seem to need an activation record, but if we want \texttt{call} to work more generally speaking, then it needs to setup an activation record.
+I have changed the descrition of \texttt{call}, so it now uses an activation record.
 }
@@
 \dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}
+
 \begin{description}
-\item[\texttt{call f(r)}] Moves the argument to a register (if there are more arguments, then they are moved to the appropriate registers).
-Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return).
-Erases all registers but the argument register(s), $\pcreg$, the $\malloc$ register (if there is one), and the return register.
-Jumps to $\texttt{f}$.
+\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}})$ are lists of registers. An overview of this call:
+  \begin{itemize}
+  \item Set up activation record
+  \item Create local enter capability for activation
+  \item Clear unused registers
+  \item Jump
+  \item Upon return: Run activation code
+  \end{itemize}
+A more detailed description of each of the above steps:
 \begin{description}
-  \item[return capability] The return capability is an enter capability into caller code.
-Upon return, the caller may assume that any return value is stored in $r_1$.
-The caller is responsible for clearing the remaining registers as part of its program.
+  \item [Set up activation record]
+    \begin{itemize}
+    \item Run malloc to get a piece of memory with space for:
+    \begin{itemize}
+      \item Words in $\bar{r}_{\var{priv}}$
+      \item Code return capability (opc)
+      \item Activation code
+    \end{itemize}
+    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
+    \item Adjust the current pc to point to the return address and save it to the activation record.
+    \item Write the activation code to the activation record.
+  \end{itemize}
+  \item [Create local enter capability for activation] Make the capability for the activaion record point to the beginning of the activation record and restrict it to a local-enter capability.
+Place this capability in $r_0$.
+  \item [Clear unused registers]
+    Clear all the register that are not $\pcreg$, $r_0$, or in $\bar{r}_{\var{args}}$.
+  \item [Jump] jump to $f$
+\lau{This needs some thought - how do I jump to $f$.
+In our semantics, you need a capability in a register in order to do a jump.
+The only way to still have the $f$ capability in a register is if it is one of the arguments.}
+  \item [Activation code] The activation code does the following:
+    \begin{itemize}
+    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
+    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
+    \item Load the return address to $\pcreg$
+    \end{itemize}
 \end{description}
@@
 Push the code that restores the stackpointer, $\pcreg$ and $r_0$.
-Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?}.
+Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?} .
 Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$.

### 2016-11-21 local-capability-machine.tex
spell checking and added new description of scall

@@
 \lau{On further thought, conveniently the examples did not seem to need an activation record, but if we want \texttt{call} to work more generally speaking, then it needs to setup an activation record.
-I have changed the descrition of \texttt{call}, so it now uses an activation record.
+I have changed the description of \texttt{call}, so it now uses an activation record.
 }
@@
 \begin{description}
-\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}})$ are lists of registers. An overview of this call:
+\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
   \begin{itemize}
@@
   \end{itemize}
-  \item [Create local enter capability for activation] Make the capability for the activaion record point to the beginning of the activation record and restrict it to a local-enter capability.
+  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability.
 Place this capability in $r_0$.
@@
 The only way to still have the $f$ capability in a register is if it is one of the arguments.}
+\lau{Maybe instead of having $f$, we use $r_f$ as the register where the ``function'' capability resides.}
   \item [Activation code] The activation code does the following:
@@
 \item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
-\item[\texttt{pop r}] Pops the top word of the stack by loading it to regiser $r$, moves 0 to the address, and decrementing the stack pointer.
-\item[\texttt{scall f(r)}] This call assumes the presence of a stack!
-Push the code that restores the stackpointer, $\pcreg$ and $r_0$.
+\item[\texttt{pop r}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
+\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
+This call assumes $r_{\var{stk}}$ contains a stack capability.
+An overview of this call:
+  \begin{itemize}
+  \item Push restore code to the stack.
+  \item Push return address capability 
+  \item Push ``private'' registers to the stack.
+  \item Push stack pointer
+  \item Create local enter capability for restoration
+  \item Restrict stack capability
+  \item Clear unused registers
+  \item Jump
+  \item Upon return: Run restore code
+  \item Return address:  Restore ``private'' state
+  \end{itemize}
+A more detailed description of the above steps:
+\begin{description}
+  \item [Push restore code to the stack]
+    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
+  \item [Push ``private'' registers to the stack]
+    Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
+  \item [Push return address capability]
+    Push a capability for the return address (on the heap) to the stack.
+  \item [Push stack pointer]
+    Push the full stackpointer to the heap.
+  \item [Create local enter capability for restoration]
+    Make a new version of the stack pointer that points to the beginning of the restoration code. Restrict it to a local-enter capability and put it in $r_0$.
+  \item [Restrict stack capability]
+    Make the stack capability only govern the unused part.
+  \item [Clear unused registers]
+    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
+  \item [Jump] Jump to $f$.
+\lau{same issue as with \texttt{call}, where dos $f$ come from.}
+  \item [Restore code]
+    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
+  \item [Return address: Restore ``private'' state]
+    \begin{itemize}
+    \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
+    \item Pop the restore code of the stack
+    \end{itemize}
+\end{description}
+
+This call assumes the presence of a stack!
+Push the code that restores the stack pointer, $\pcreg$ and $r_0$.
 Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?} .
 Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$.
-Restrict the stackpointer to the ``unused part''.
+Restrict the stack pointer to the ``unused part''.
 Push the arguments to the stack and jump to \texttt{f}.

### 2016-11-21 local-capability-machine.tex
removed figures (for now)

@@
 \usepackage{mathpartir}
+
+\usepackage{rotating}
 
@@
 \end{description}
-
-This call assumes the presence of a stack!
-Push the code that restores the stack pointer, $\pcreg$ and $r_0$.
-Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?} .
-Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$.
-Restrict the stack pointer to the ``unused part''.
-Push the arguments to the stack and jump to \texttt{f}.
-  \begin{description}
-  \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code.
-\dominique{restores other registers too, right?}
 \end{description}
-Figure \ref{fig:stack-before-call}-\ref{fig:stack-after-call} illustrates what the stack would look like at different stages of the scall.
-\end{description}
-
+\begin{comment}
+\lau{todo: adjust figures}
 \begin{figure}
@@
 \cline{2-2}
-$r_{\var{stk}} \rightarrow$   & local state\\
+$r_{\var{stk}} \rightarrow$   & local stack\\
    & $\vdots$
+\end{tabular}
+\hspace{0.5cm}
+\begin{tabular}{*{10}{>{\centering\arraybackslash}p{0.2cm}}}
+  \begin{sideways}
+    $\pcreg$
+  \end{sideways}
+& 
+  \begin{sideways}
+    $r_{\var{stk}}$
+  \end{sideways}
+& 
+  \begin{sideways}
+    $r_0$
+  \end{sideways}
+& 
+  \begin{sideways}
+    $r_{\var{args},1}$
+  \end{sideways}
+&
+
+& 
+  \begin{sideways}
+    $r_{\var{args},n}$ 
+  \end{sideways}
+& 
+  \begin{sideways}
+    $r_{\var{priv},1}$
+  \end{sideways}
+&
+
+&
+  \begin{sideways}
+    $r_{\var{priv},n}$
+  \end{sideways}
+&\\
+\hline
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$\pc$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$c_s$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$c_0$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$w_1$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+ $\phantom{1}$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$w_n$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$v_1$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+  $\phantom{1}$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm}}{
+$v_n$
+} &
+\multicolumn{1}{|>{\centering\arraybackslash}p{0.2cm} | }{
+  $\phantom{1}$
+}\\
+\hline
 \end{tabular}
@@
 \end{figure}
+\end{comment}
 \subsection{Labels}

### 2016-11-22 local-capability-machine.tex
Added figures to illustrate scall

--- 2016-11-21 local-capability-machine.tex
@@
   \item Push restore code to the stack.
+  \item Push ``private'' registers to the stack.
   \item Push return address capability 
-  \item Push ``private'' registers to the stack.
   \item Push stack pointer
@@
   \item Return address:  Restore ``private'' state
+\lau{ Need to clear the returned stack. }
   \end{itemize}
@@
 \end{description}
-
-This call assumes the presence of a stack!
-Push the code that restores the stack pointer, $\pcreg$ and $r_0$.
-Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?} .
-Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$.
-Restrict the stack pointer to the ``unused part''.
-Push the arguments to the stack and jump to \texttt{f}.
-  \begin{description}
-  \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code.
-\dominique{restores other registers too, right?}
-  \end{description}
-Figure \ref{fig:stack-before-call}-\ref{fig:stack-after-call} illustrates what the stack would look like at different stages of the scall.
 \end{description}
@@
   \centering
-  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Heap} \\
    & \\
@@
 \cline{2-2}
-$r_{\var{stk}} \rightarrow$   & local state\\
+$c_{\var{stk}} \rightarrow$   & local stack\\
    & $\vdots$
 \end{tabular}
-\caption{Stack after everything stored}
+\hspace{1cm}
+\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $c_{\pc}$\\
+\cline{2-2}
+$r_0$  & $c_0$ \\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}$ \\
+\cline{2-2}
+$r_{\var{args},1}$ & $w_{a,1}$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{args},n}$ & $w_{a,1}$\\
+\cline{2-2}
+$r_{\var{priv},1}$ & $w_{p,1}$\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{priv},m}$ & $w_{p,m}$\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{Stack and register-file before the call}
 \end{figure}
 
+\begin{figure}
+  \label{fig:stack-after-push}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Heap} \\
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+$c_{\var{stk}}' \rightarrow$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+   & $c_\pc'$ \\
+\cline{2-2}
+   & $c_0$ \\
+\cline{2-2}
+   & $w_{p,1}$ \\
+\cline{2-2}
+   & $\vdots$ \\
+\cline{2-2}
+   & $w_{p,m}$ \\
+\cline{2-2}
+   & Restore code \\
+\cline{2-2}
+   & local stack\\
+   & $\vdots$
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $c_{\pc}''$\\
+\cline{2-2}
+$r_0$  & $c_0$ \\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+$r_{\var{args},1}$ & $w_{a,1}$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{args},n}$ & $w_{a,1}$\\
+\cline{2-2}
+$r_{\var{priv},1}$ & $w_{p,1}$\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{priv},m}$ & $w_{p,m}$\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{Stack and register-file after everything has been pushed to the stack.}
+\end{figure}
+
+\begin{figure}
+  \label{fig:stack-after-restrict-and-zero}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Heap} \\
+   & \\
+   & $\vdots$\\
+\cline{2-2}
+$c_{\var{stk}}'' \rightarrow$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+   & $c_\pc'$ \\
+\cline{2-2}
+   & $c_0$ \\
+\cline{2-2}
+   & $w_{p,1}$ \\
+\cline{2-2}
+   & $\vdots$ \\
+\cline{2-2}
+   & $w_{p,m}$ \\
+\cline{2-2}
+$c_0' \rightarrow$   & Restore code \\
+   & $\vdots$\\
+\cline{2-2}
+   & local stack\\
+   & $\vdots$
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $c_{\pc}^{(3)}$\\
+\cline{2-2}
+$r_0$  & $c_0'$ \\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}''$ \\
+\cline{2-2}
+$r_{\var{args},1}$ & $w_{a,1}$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{args},n}$ & $w_{a,1}$\\
+\cline{2-2}
+$r_{\var{priv},1}$ & 0\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{priv},m}$ & 0 \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{ Stack and register-file after the $c_{\var{stk}}'$ has been limited to only give authority over the empty part of the stack (the new capability is $c_{\var{stk}}''$.
+$c_0'$ is made from $c_{\var{stk}}'$ by setting it to point to the restore code and restricting it to a local enter-capability.
+}
+\end{figure}
+
+\begin{figure}
+  \label{fig:stack-upon-return}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Heap} \\
+   & \\
+   & $\vdots$\\
+\cline{2-2} 
+   & $?$\\
+\cline{2-2} 
+$c_{\var{stk}}' \rightarrow$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+   & $c_\pc'$ \\
+\cline{2-2}
+   & $c_0$ \\
+\cline{2-2}
+   & $w_{p,1}$ \\
+\cline{2-2}
+   & $\vdots$ \\
+\cline{2-2}
+   & $w_{p,m}$ \\
+\cline{2-2}
+$c_0' \rightarrow$   & Restore code \\
+   & $\vdots$\\
+\cline{2-2}
+   & local stack\\
+   & $\vdots$
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $c_0'$\\
+\cline{2-2}
+$r_0$  &  ? \\
+\cline{2-2}
+$r_{\var{stk}}$  & ? \\
+\cline{2-2}
+$r_1$ & $w_1$ \\
+\cline{2-2}
+$r_{\var{priv},1}$ & ?\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{priv},m}$ & ? \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{ Stack and register-file upon return. }
+\end{figure}
+
+\begin{figure}
+  \label{fig:stack-after-restore-code}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Heap} \\
+   & \\
+   & $\vdots$\\
+\cline{2-2} 
+   & $?$\\
+\cline{2-2} 
+   & 0 \\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
+$c_{\var{stk}}^{(3)} \rightarrow$  & $c_0$ \\
+\cline{2-2}
+   & $w_{p,1}$ \\
+\cline{2-2}
+   & $\vdots$ \\
+\cline{2-2}
+   & $w_{p,m}$ \\
+\cline{2-2}
+$c_0' \rightarrow$   & Restore code \\
+   & $\vdots$\\
+\cline{2-2}
+   & local stack\\
+   & $\vdots$
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $c_\pc'$\\
+\cline{2-2}
+$r_0$  &  ? \\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}^{(3)}$ \\
+\cline{2-2}
+$r_1$ & $w_1$ \\
+\cline{2-2}
+$r_{\var{priv},1}$ & ?\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{priv},m}$ & ? \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{ Stack and register-file upon return. }
+\end{figure}
+
+\begin{figure}
+  \label{fig:stack-after-restore-code}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Heap} \\
+   & \\
+   & $\vdots$\\
+\cline{2-2} 
+   & $?$\\
+\cline{2-2} 
+   & 0 \\
+\cline{2-2}
+   & $\vdots$\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
+$c_{\var{stk}}^{(4)} \rightarrow$   & local stack\\
+   & $\vdots$
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $c_\pc^{(3)}$\\
+\cline{2-2}
+$r_0$  &  $c_0$ \\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}^{(4)}$ \\
+\cline{2-2}
+$r_1$ & $w_1$ \\
+\cline{2-2}
+$r_{\var{priv},1}$ & $w_{p,1}$\\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+$r_{\var{priv},m}$ & $w_{p,m}$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{ Stack and register-file upon return. }
+\end{figure}
+
+\begin{comment}
 \begin{figure}

### 2016-11-22 local-capability-machine.tex
removed old figures
Adjusted scall and call semantics to use explicit register

@@
 \begin{description}
-\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
+\item[\texttt{call} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
   \begin{itemize}
@@
   \item [Clear unused registers]
-    Clear all the register that are not $\pcreg$, $r_0$, or in $\bar{r}_{\var{args}}$.
-  \item [Jump] jump to $f$
-\lau{This needs some thought - how do I jump to $f$.
-In our semantics, you need a capability in a register in order to do a jump.
-The only way to still have the $f$ capability in a register is if it is one of the arguments.}
-\lau{Maybe instead of having $f$, we use $r_f$ as the register where the ``function'' capability resides.}
+    Clear all the register that are not $\pcreg$, $r_0$, $r_f$ or in $\bar{r}_{\var{args}}$.
+  \item [Jump] jump to $r_f$
   \item [Activation code] The activation code does the following:
@@
 \item[\texttt{pop r}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
-\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
+\item[\texttt{scall} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
 This call assumes $r_{\var{stk}}$ contains a stack capability.
@@
   \item Upon return: Run restore code
-  \item Return address:  Restore ``private'' state
-\lau{ Need to clear the returned stack. }
+  \item Return address: Clear the stack segment we gave control over
+  \item Restore ``private'' state
   \end{itemize}
@@
     Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
-  \item [Return address: Restore ``private'' state]
+  \item [Return Address:] \forcenewline
+    \begin{description}
+    \item [Clear the stack segment we gave control over]
+    \item [Restore ``private'' state] \forcenewline
       \begin{itemize}
@@
 \end{description}
+\end{description}
 
@@
   \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
-\multicolumn{2}{l}{Heap} \\
+\multicolumn{2}{l}{Stack} \\
    & \\
@@
 \cline{2-2}
+$r_f$ & $c_f$ \\
+\cline{2-2}
 $r_{\var{args},1}$ & $w_{a,1}$ \\
@@
   \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
-\multicolumn{2}{l}{Heap} \\
+\multicolumn{2}{l}{Stack} \\
+\cline{2-2}
    & \\
@@
    & local stack\\
-   & $\vdots$
+   & $\vdots$ \\
+\cline{2-2}
 \end{tabular}
@@
    & local stack\\
-   & $\vdots$
+   & $\vdots$\\
+\cline{2-2}
 \end{tabular}
@@
 \cline{2-2}
-   & $?$\\
-\cline{2-2} 
    & 0 \\
 \cline{2-2}
-   & 0 \\
-\cline{2-2}
 $c_{\var{stk}}^{(3)} \rightarrow$  & $c_0$ \\
@@
 \cline{2-2}
-$c_0' \rightarrow$   & Restore code \\
+   & Restore code \\
    & $\vdots$\\
@@
 \end{tabular}
-\caption{ Stack and register-file upon return. }
+\caption{ Stack and register-file after clearing stack. }
 \end{figure}
@@
 \cline{2-2}
-   & $?$\\
-\cline{2-2} 
    & 0 \\
 \cline{2-2}
+$c_{\var{stk}}^{(4)} \rightarrow$   & local stack\\
    & $\vdots$\\
 \cline{2-2}
-   & 0 \\
-\cline{2-2}
-$c_{\var{stk}}^{(4)} \rightarrow$   & local stack\\
-   & $\vdots$
 \end{tabular}
@@
 
-\begin{comment}
-\begin{figure}
-  \centering
-  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
-   & \\
-   & $\vdots$\\
-\cline{2-2}
-   & 0 \\
-\cline{2-2}
-$r_{\var{stk}} \rightarrow$     & Stack cap\\
-\cline{2-2}
-   & $r_0$-continuation\\
-\cline{2-2}
-   & o-$\pcreg$\\
-\cline{2-2}
-   & Restore code\\
-   & $\vdots$\\
-\cline{2-2}
-   & local state\\
-   & $\vdots$
-\end{tabular}
-\caption{Stack after everything stored}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
-   & \\
-   & $\vdots$\\
-\cline{2-2}
-   & 0 \\
-\cline{2-2}
-$r_{\var{stk}} \rightarrow$      & arg1 \\
-\cline{2-2}
-   & $\vdots$\\
-\cline{2-2}
-   & argn \\
-\cline{2-2}
-   & Stack cap\\
-\cline{2-2}
-   & $r_0$-continuation\\
-\cline{2-2}
-   & o-$\pcreg$\\
-\cline{2-2}
-   & Restore code\\
-   & $\vdots$\\
-\cline{2-2}
-   & local state\\
-   & $\vdots$
-\end{tabular}
-\caption{Stack after arguments pushed to stack an ready to jump}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{tabular}[!h]{r|>{\centering\arraybackslash}p{3cm}|l}
-   & &\\
-   & $\vdots$ &\\
-\cline{2-2}
-   & 0 &\\
-\cline{2-2}
-$r_{\var{stk}} \rightarrow$      & arg1 & \\
-\cline{2-2}
-   & $\vdots$ & \\
-\cline{2-2}
-   & argn & \\
-\cline{2-2}
-   & Stack cap & \\
-\cline{2-2}
-   & $r_0$-continuation &  \\
-\cline{2-2}
-   & o-$\pcreg$ & \\
-\cline{2-2}
-   & Restore code & \rdelim\}{2}{2cm}[variable length]\\
-   & $\smash{\vdots}$ & \\
-\cline{2-2}
-   & & \\
-\cline{2-2}
-   & local state & \\
-   & $\smash{\vdots}$ &
-\end{tabular}
-\caption{Stack after arguments pushed to stack an ready to jump}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
-   & \\
-   & $\vdots$\\
-\cline{2-2}
-   & ? \\
-\cline{2-2}
-   & Stack cap\\
-\cline{2-2}
-   & $r_0$-continuation\\
-\cline{2-2}
-   & o-$\pcreg$\\
-\cline{2-2}
-$\pcreg \rightarrow$   & Restore code\\
-   & $\vdots$\\
-\cline{2-2}
-   & local state\\
-   & $\vdots$
-\end{tabular}
-\caption{Stack just after return}
-\end{figure}
-
-\begin{figure}
-  \label{fig:stack-after-call}
-  \centering
-  \begin{tabular}[!h]{r | >{\centering\arraybackslash}p{3cm} |}
-   & \\
-   & $\vdots$\\
-\cline{2-2}
-   & 0 \\
-\cline{2-2}
-$r_{\var{stk}} \rightarrow$   & local state\\
-   & $\vdots$
-\end{tabular}
-\caption{Stack after everything has been restored}
-\end{figure}
-\end{comment}
 \subsection{Labels}

### 2016-11-22 local-capability-machine.tex
Addressed Dominique's comments in f1 lemma
removed comment and fixed call jump

@@
   \item [Clear unused registers]
-    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
-  \item [Jump] Jump to $f$.
-\lau{same issue as with \texttt{call}, where dos $f$ come from.}
+    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, $r_f$, and $\bar{r}_{\var{args}}$.
+  \item [Jump] Jump to $r_f$.
   \item [Restore code]
@@
   \end{align*}
-if \dominique{$\hs_f$ \emph{and} $\hs_{f1}$?}
+if 
   \[
     (\reg_0\update{\pcreg}{c_{\adv}}
-          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_f) \step[n] (\halted,h')
+          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_{\var{frame}}) \step[n] (\halted,h')
   \]
@@
 \end{align*}
-and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ \dominique{do you mean f1?} including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
+and $hs_\flag(\flag) = 0$. $hs_{f1}$ contains the code of $f1$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
 then

### 2016-11-22 local-capability-machine.tex
Added comments about checks for the interesting example.

@@
                fun f =>
-                 x:=0;
+                 x := 0;
                  f();
-                 x:=1;
+                 x := 1;
                  f();
-                 assert(x==1)
+                 assert(x == 1)
 \end{verbatim}
-\lau{ This example relies on \texttt{adv} actually passing on a proper $\malloc$ - is that a fair assumption?
-The example can be changed to not require this by assuming x and the function is already in the heap.
-This change does not remove the difficulties in the original program.
+
+\lau{In order for this to be secure using a stack the closure needs to make a couple of checks: 1) The alleged stack capability has to be local and $\rwlx$ - if not, then fail.
+2) The callback \texttt{f} has to be a global capability - if not, then fail.
+3) The stack needs to be cleared before the first callback.
+\\If it looks like a stack, works like a stack, and quacks like a stack, then it is probably a stack.
+The first requirement makes sure that what was passed as the stack can be used as a stack eventhough it might not be a stack capability that follows the conventional stack diciplin.
+If the capability is actually for part of the callers stack, then we will just overwrite their part of the stack.
+We are not guaranteed that the caller did not save the stack capability - they did after all have a stack to save it on.
+In fact, if the caller wants to reuse our part of the stack later, then they have to keep a capability for our part of the stack around.
+When the caller is not trusted, then this seems like an unsafe practice.
+This brings us to the second requirement: If the callback is global, then there is no way it can restore the previous stack capability and gain access to our part of the stack.
+This is the case as the \emph{only} place they could have stored the stack capability is on the stack, and the only way they can restore this is through one of the local stack capabilities.
+In other words, this requirement makes sure that the callback cabaility is not derived from the stack pointer (which would allow them to get a stack capability for our part of the stack).\\ The last requirement is to sanitize the stack.
+This makes sure that the caller does not try to sneak a capability to the callback simply by leaving it somewhere on the stack and hoping that we won't notice and pass it on.
 }
+
 \begin{lemma}\forcenewline

### 2016-11-22 local-capability-machine.tex
Added size to malloc instruction

@@
     \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
-    \item Adjust the current pc to point to the return address and save it to the activation record.
+    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
     \item Write the activation code to the activation record.
@@
     \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
-    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
+%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
     \item Load the return address to $\pcreg$
@@
 \end{description}
-\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
-\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
-\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them.
+\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
+\item[\texttt{store $r$ $n$}] Stores $n$ to $r$ - notice the semantics do not allow an integer to be stored directly.
+\item[\texttt{assert $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$.
 If they are equal, then execution continues.
@@
 \end{description}
-
+Note:
+\begin{itemize}
+\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local.
+Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
+\end{itemize}
 \subsection{Stack}
@@
 \begin{description}
-\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
-\item[\texttt{pop r}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
+\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
+\item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
 \item[\texttt{scall} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
@@
   \item [Push stack pointer]
-    Push the full stackpointer to the heap.
+    Push the full stackpointer to the stack.
   \item [Create local enter capability for restoration]
@@
     \begin{description}
-    \item [Clear the stack segment we gave control over]
+    \item [Clear the stack segment we gave control over] Store 0 to all the cells the stack capability passed to the callee governed.
     \item [Restore ``private'' state] \forcenewline
@@
 \end{description}
-
+Note:
+\begin{itemize}
+\item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared upon return.
+\item As a rule of thumb: If you receive a stack from an untrusted source, then you need to check that it is a local $\rwlx$-capability and clear it!
+If any callbacks are provided, then they need to be global.
+\end{itemize}
 \begin{figure}

### 2016-11-22 local-capability-machine.tex
Added clear macro instruction
Added note on the use of call, scall and the use of a combination of the two

@@
 \end{description}
+\item[\texttt{clear $r$}] Clears all cells $r$ has authority over. 
 \end{description}
@@
 \begin{itemize}
+\item If we want to have local capabilities as part of our private state, then we need to have a stack and use \texttt{scall}.
+If we do not have any local capabilities we want to keep around, then we can use \texttt{call}, but it will incur a small memory leak as the activation records cannot be recycled!
+It is also possible to use a combination of \texttt{scall} and \texttt{call}, but when \texttt{call} is used, then we have no way to store the stack, so we cannot use \texttt{scall} after that.
 \item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared upon return.

### 2016-11-22 local-capability-machine.tex
Comment

@@
   \item Push stack pointer
-  \item Create local enter capability for restoration
+  \item Create protected return pointer
   \item Restrict stack capability
@@
     Push the full stackpointer to the stack.
-  \item [Create local enter capability for restoration]
-    Make a new version of the stack pointer that points to the beginning of the restoration code. Restrict it to a local-enter capability and put it in $r_0$.
+  \item [Create protected return pointer]
+    Make a new version of the stack pointer that points to the beginning of the restoration code.
+Restrict it to a local enter-capability and put it in $r_0$.
+\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?}
   \item [Restrict stack capability]

### 2016-11-22 local-capability-machine.tex
Added lemmas about n-equality and future worlds
Added linking list and flag list. Changed terminologi, so what before was heap now is memory. Any occurence of heap in the pdf is now most likely a mistake

@@
 \newcommand{\var}[1]{\mathit{#1}}
-\newcommand{\hs}{\var{hs}}
+\newcommand{\hs}{\var{ms}}
+\newcommand{\ms}{\ms}
 \newcommand{\hv}{\var{hv}}
@@
 \newcommand{\reg}{\var{reg}}
-\newcommand{\heapseg}{\var{hs}}
-\newcommand{\heap}{\var{heap}}
+\newcommand{\heapseg}{\var{ms}}
+\newcommand{\heap}{\var{mem}}
 \newcommand{\mode}{\var{mode}}
@@
 \newcommand{\plainproj}[1]{\mathrm{#1}}
-\newcommand{\memheap}[1][\Phi]{#1.\plainproj{heap}}
+\newcommand{\memheap}[1][\Phi]{#1.\plainproj{mem}}
 \newcommand{\memreg}[1][\Phi]{#1.\plainproj{reg}}
 
-\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{heap}.#2}{#3}}
+\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{mem}.#2}{#3}}
 \newcommand{\updateReg}[3][\Phi]{#1\update{\plainproj{reg}.#2}{#3}}
@@
 \newcommand{\Addrs}{\plaindom{Addr}}
-\newcommand{\Mems}{\plaindom{Mem}}
+\newcommand{\ExecConfs}{\plaindom{ExecConf}}
 \newcommand{\RegName}{\plaindom{RegisterName}}
 \newcommand{\Regs}{\plaindom{Reg}}
-\newcommand{\Heaps}{\plaindom{Heap}}
-\newcommand{\HeapSegments}{\plaindom{HeapSegment}}
+\newcommand{\Heaps}{\plaindom{Mem}}
+\newcommand{\HeapSegments}{\plaindom{MemSegment}}
 \newcommand{\Confs}{\plaindom{Conf}}
@@
 \newcommand{\xistr}{\xi_{\var{str}}}
-\newcommand{\StorePred}{\plaindom{HeapSegPred}}
+\newcommand{\StorePred}{\plaindom{MemSegPred}}
 \newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
@@
 
-%Reference register/heap
+%Reference register/memory
 \newcommand{\refreg}[1]{\lfloor #1 \rfloor}
-\newcommand{\refheap}[1]{\langle #1 \rangle_h}
+\newcommand{\refheap}[1]{\langle #1 \rangle_m}
 
@@
 \begin{align*}
-\Addrs &::= \nats & & &
+\Addrs &::= \nats\\
 \Words &::= \Caps + \ints \\
-\Regs  &::= \RegName \rightarrow \Words & & &
+\Regs  &::= \RegName \rightarrow \Words\\
 \Heaps &::= \Addrs \rightarrow \Words \\
-\Perms &::= \{ \noperm, \readonly, \readwrite, \readwritel, \exec, \entry, \rwx, \rwlx\} & & &
-\Mems  &::= \Regs \times \Heaps \\
-\Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs & & &
-\Confs &::= \Mems + \{\failed, \halted \times \Mems\} \\
-\HeapSegments &::= \Addrs \parfun \Words & & &
+\Perms &::= \{ \noperm, \readonly, \readwrite, \readwritel, \exec, \entry, \rwx, \rwlx\}\\
+\ExecConfs  &::= \Regs \times \Heaps \\
+\Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
+\Confs &::= \ExecConfs + \{\failed, \halted \times \ExecConfs\} \\
+\HeapSegments &::= \Addrs \parfun \Words\\
 \Globals & ::= \{\glob, \local \}
@@
 r       &\in& \RegName\\
-\mem    &::=& (\reg,\heap)\\
 \pc     &\in& \Caps \\
 \pcreg  &\in& \RegName \\
-\Phi    &::=& \mem \in \Confs\\
+\Phi    &\in& \ExecConfs \\
 \memheap&\in& \Heaps \\
@@
 \perm   &\in& \Perms\\
-(\perm,\start,\addrend,\addr) &\in& \Caps \\
+((\perm,\gl),\start,\addrend,\addr) &\in& \Caps \\
 n       &\in& \ints\\
@@
   \begin{align*}
-    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
+    &\forall \Phi \in \ExecConfs \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
     &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
@@
     &\qquad \quad \Rightarrow \\
-    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
+    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
     &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
@@
 \end{specification}
+\lau{Only configurations with a memory (not memory segments) are executable.
+Therefore either $\hsfoot$ or $\hsframe$ need to be a memory and not a memory segment (it should probably be the footprint?).}
 
@@
 
+\subsection{Linking and ABI}
+In order to make capabilities to trusted code (and possibly untrusted code) available, we assume that some sort of linker has made these available.
+This is done in the following way: For every function, the first memory cell the capability for that function governs contains a capability for the list in question.
+Each function name in a program corresponds to an offset in the table, e.g., \texttt{malloc}.
+When a name is used in a program, it indicates what entry to pick from the list.
+The list should always be accessible by taking a copy of the capability in the $\pcreg$-register and adjusting it to point to the first cell it governs.
+
+The capability link list can be shared between multiple functions as it is accessed through a read-only capability.
+
+\subsection{Flag list}
+A function may use flags to signal failure.
+We use the convention that a flag list is available in the second memory cell of a functions code (so just after the linking table).
+The flag list is accessed through a read-write capability and initially it contains all zero.
+Like the linking list, each entry is associated with a name which may appear in the macros.
+
+The flag list should never be shared between distrusting parties.
+
 \begin{description}
-\item[\texttt{call} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
+\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
   \begin{itemize}
@@
     Clear all the register that are not $\pcreg$, $r_0$, $r_f$ or in $\bar{r}_{\var{args}}$.
-  \item [Jump] jump to $r_f$
+  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
   \item [Activation code] The activation code does the following:
@@
 \item[\texttt{store $r$ $n$}] Stores $n$ to $r$ - notice the semantics do not allow an integer to be stored directly.
-\item[\texttt{assert $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$.
+\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$.
 If they are equal, then execution continues.
-If they are unequal, then the assertion flag is set to 1 and execution halts.
+If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
 \end{description}
@@
 Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
+
+The stack will reside entirely in memory. There is no seperation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
+
 \begin{description}
@@
 \item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
-\item[\texttt{scall} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
+\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
 This call assumes $r_{\var{stk}}$ contains a stack capability.
@@
   \item [Push return address capability]
-    Push a capability for the return address (on the heap) to the stack.
+    Push a capability for the return address (in the memory) to the stack.
   \item [Push stack pointer]
@@
     Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, $r_f$, and $\bar{r}_{\var{args}}$.
-  \item [Jump] Jump to $r_f$.
+  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
   \item [Restore code]
@@
     (\reg_0\update{\pcreg}{c_{\adv}}
-          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_{\var{frame}}) \step[n] (\halted,h')
+          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus m_{\malloc} \uplus \hs_{\var{frame}}) \step[n] (\halted,m')
   \]
@@
 \end{align*}
-and $hs_\flag(\flag) = 0$. $hs_{f1}$ contains the code of $f1$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
+and $\hs_\flag(\flag) = 0$. $\hs_{f1}$ contains the code of $f1$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
 then
 \[
-  h'(\flag) = 0
+  m'(\flag) = 0
 \]  
@@
   \[
-    (\reg_0\update{\pcreg}{c_f}\update{r_1}{c_\adv},\hs_\adv \uplus \hs_f \uplus \hs_\malloc \uplus h_\flag \uplus h_\var{frame}) \step[n] (\halted, h')
+    (\reg_0\update{\pcreg}{c_f}\update{r_1}{c_\adv},\hs_\adv \uplus \hs_f \uplus m_\malloc \uplus \hs_\flag \uplus \hs_{\var{frame}}) \step[n] (\halted, h')
   \]
@@
     (\reg_0\update{\pcreg}{c_{\adv}}
-          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus h_f) \step[n] (\halted,h')
+          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus m_f) \step[n] (\halted,m')
   \]
+\lau{the frame should probably just be a segment}
 where 
@@
 \[
-  h'(\flag) = 0
+  m'(\flag) = 0
 \]
@@
 
-Heap satisfaction:
+Memory segment satisfaction:
 \begin{align*}
@@
 We expect the following lemmas to hold true:
-\subsubsection{Heap Satisfaction}
+\subsubsection{Memory Segment Satisfaction}
 \begin{lemma}[Downwards closure]
@@
   \end{itemize}
+\end{lemma}
+
+\begin{lemma}
+  \begin{align*}
+    &\forall n,W_1,W_2,W_1' \ldotp
+    &\quad W_1 \nequal W_2 \land W_1' \futurewk W_1 \Rightarrow \exists W_2' \ldotp W_2' \nequal W_1' \land W_2' \futurewk W_2
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}
+  \begin{align*}
+    &\forall n,W_1,W_2,W_1' \ldotp
+    &\quad W_1 \nequal W_2 \land W_1' \futurestr W_1 \Rightarrow \exists W_2' \ldotp W_2' \nequal W_1' \land W_2' \futurestr W_2
+  \end{align*}
 \end{lemma}

### 2016-11-23 local-capability-machine.tex
Added text to the figures that illustrate scall

@@
   \item [Clear unused registers]
-    Clear all the register that are not $\pcreg$, $r_0$, $r_f$ or in $\bar{r}_{\var{args}}$.
+    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
   \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
@@
 \item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
-\item[\texttt{store $r$ $n$}] Stores $n$ to $r$ - notice the semantics do not allow an integer to be stored directly.
 \item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$.
@@
 The stack will reside entirely in memory. There is no seperation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
+
+Eventhough the memory is infinite, we will only use a finite part for the stack.
+If we have allocated too little memory for the stack, and we try to push something anyway, then the execution will fail.
+As we consider failing admissable, we are okay with this.
 
@@
   \item [Clear unused registers]
-    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, $r_f$, and $\bar{r}_{\var{args}}$.
+    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
   \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
@@
 \end{description}
-\item[\texttt{clear $r$}] Clears all cells $r$ has authority over. 
+\item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 
 \end{description}
@@
 \cline{2-2}
-$r_f$ & $c_f$ \\
-\cline{2-2}
 $r_{\var{args},1}$ & $w_{a,1}$ \\
@@
 \cline{2-2}
-$r_{\var{args},n}$ & $w_{a,1}$\\
+$r_{\var{args},n}$ & $w_{a,n}$\\
 \cline{2-2}
@@
 \end{tabular}
-\caption{Stack and register-file before the call}
+\caption{This is the first figure of 6 that illustrates how \texttt{scall} works.
+In this example, the call \texttt{scall $f([r_{\var{args},1},\dots,r_{\var{args},n}],[r_0,r_{\var{priv},1},\dots,r_{\var{priv},m}])$.
+In this example the two lists of registers are disjoined eventhough that does not have to be the case.}}
 \end{figure}
@@
 \end{tabular}
-\caption{Stack and register-file after everything has been pushed to the stack.}
+\caption{Stack and register-file after the restore code, ``private'' registers (remember $r_0$ is here private.), return address ($c_\pc'$), and stack capability ($c_{\var{stk}}'$) have been pushed to the stack.}
 \end{figure}
@@
 \end{tabular}
-\caption{ Stack and register-file after the $c_{\var{stk}}'$ has been limited to only give authority over the empty part of the stack (the new capability is $c_{\var{stk}}''$.
+\caption{ Stack and register-file after the $c_{\var{stk}}'$ has been limited to only give authority over the empty part of the stack (the new capability is $c_{\var{stk}}''$).
 $c_0'$ is made from $c_{\var{stk}}'$ by setting it to point to the restore code and restricting it to a local enter-capability.
-}
+The ``private'' registers have been cleared.}
 \end{figure}
@@
 \end{tabular}
-\caption{ Stack and register-file upon return. }
+\caption{ Stack and register-file upon return from $f$.
+At this point we have no idea what is in the register-file apart from the $\pcreg$ which we know points to the restore code.
+The contents of the stack we released access to is also unknown.
+(Notice that we have changed the order of the registers as we are no longer interested in the argument registers.
+By convention we expect a return value to be in $r_1$, which is why we have named that word, but the words in the remaining non-special-purpose registers could also be considered return values.)}
 \end{figure}
@@
 \cline{2-2}
-   & 0 \\
+   & ? \\
 \cline{2-2}
@@
 \end{tabular}
-\caption{ Stack and register-file after clearing stack. }
+\caption{ Stack and register-file after executing the restore code. The old stack capability has been restored and the  $\pcreg$-register now points to the return address in memory. }
 \end{figure}
@@
 \cline{2-2}
-$c_{\var{stk}}^{(4)} \rightarrow$   & local stack\\
+$c_{\var{stk}} \rightarrow$   & local stack\\
    & $\vdots$\\
@@
 \cline{2-2}
-$r_{\var{stk}}$  & $c_{\var{stk}}^{(4)}$ \\
+$r_{\var{stk}}$  & $c_{\var{stk}}$ \\
 \cline{2-2}
@@
 \cline{2-2}
+& $?$ \\
+\cline{2-2}
 & $\vdots$ \\
@@
 \end{tabular}
-\caption{ Stack and register-file upon return. }
+\caption{ Stack and register-register file after the clean up code has been run.
+The untrusted part of the stack has been cleared.
+The ``private'' words have been poped to their respective registers.
+The restore code has been poped of the stack.
+The stack is now in the same state as it was before the call.
+}
 \end{figure}

### 2016-11-23 local-capability-machine.tex
updated f1 and f2

@@
 \item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
-\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$.
+\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison).
 If they are equal, then execution continues.
@@
 \multicolumn{2}{l}{Stack} \\
+\cline{2-2}
    & \\
@@
 $c_{\var{stk}} \rightarrow$   & local stack\\
-   & $\vdots$
+   & $\vdots$\\
+\cline{2-2}
 \end{tabular}
@@
 \end{verbatim}
-Assembly program not using stack
+Assembly program not using stack. Assume that $r_l \not\in \{\pcreg,r_0 \}$ is a register.
 \begin{verbatim}
-f1: malloc l
-    hstore l 1
-    call adv()
-    assert l 1
+f1: malloc r_l 1
+    store r_l 1
+    call adv([],[r_l])
+    assert r_l 1
     halt
@@
 f2: push 1
-    scall adv()
+    scall adv([],[])
     pop r1

### 2016-11-23 local-capability-machine.tex
Updated lemma for f1

@@
 \newcommand{\hs}{\var{ms}}
-\newcommand{\ms}{\ms}
+\newcommand{\ms}{\hs}
 \newcommand{\hv}{\var{hv}}
@@
 \newcommand{\adv}{\var{adv}}
+\newcommand{\msframe}{ms_\var{frame}}
+\newcommand{\link}{\var{link}}
 \newcommand{\flag}{\var{flag}}
@@
 \newcommand{\rio}{r_{io}}
-\renewcommand{\adv}{\codelabel{adv}}
 \newcommand{\advb}{\var{adv_{base}}}
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
-    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
-  \end{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+2) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-2,\mathtt{1f},\mathtt{f1}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+2) \\
+    m & \defeq \hs_{f1} \uplus 
+               \hs_\flag \uplus                
+               \ms_{\var{link}} \uplus 
+               \hs_\adv \uplus 
+               m_{\malloc} \uplus 
+               \hs_{\var{frame}} 
+  \end{align*}
+where 
+\begin{align*}
+  &\dom(\hs_{f1}) = [\mathtt{f1}-2,\mathtt{1f}] \\
+  &\dom(\hs_\flag) = [\flag,\flag] \\
+  &\dom(\ms_\link) = [\link,\link+1]
+  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
+\end{align*}
+and
+\begin{itemize}
+\item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_{flag}$, the rest of $\hs_{f1}$ contains the code of $f1$.
+\item $\ms_\flag = [flag \mapsto 0]$
+\item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+\item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
+\end{itemize}
 if 
   \[
-    (\reg_0\update{\pcreg}{c_{\adv}}
-          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus m_{\malloc} \uplus \hs_{\var{frame}}) \step[n] (\halted,m')
+    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m')
   \]
-where 
-\begin{align*}
-  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
-  &\dom(\hs_\flag) = [\flag,\flag] \\
-  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
-\end{align*}
-and $\hs_\flag(\flag) = 0$. $\hs_{f1}$ contains the code of $f1$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
 then

### 2016-11-23 local-capability-machine.tex
Added lemma for f2

@@
 \newcommand{\link}{\var{link}}
+\newcommand{\stk}{\var{stk}}
 \newcommand{\flag}{\var{flag}}
@@
 
+When not in the middle of a push or a pop, the stack capability points to the top word of the stack.
+For an empty stack, the stack capability points to the address just outside of the range of authority for the stack capability.
+
 \begin{description}
-\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
+\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the stack capability by one and storing the word through said capability.\footnote{The stack grows upwards.}
 \item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
@@
     assert r_l 1
-    halt
-1f:
+1f: halt
+
 \end{verbatim}
-\begin{lemma}[Correctness lemma for \texttt{f1}]
+\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
   For all $n \in \nats$
@@
   &\dom(\hs_\flag) = [\flag,\flag] \\
-  &\dom(\ms_\link) = [\link,\link+1]
+  &\dom(\ms_\link) = [\link,\link+1]\\
   &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
@@
 \begin{itemize}
-\item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_{flag}$, the rest of $\hs_{f1}$ contains the code of $f1$.
+\item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f1}$ contains the code of $f1$.
 \item $\ms_\flag = [flag \mapsto 0]$
 \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-\item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
+\item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
 \end{itemize}
@@
   \[
-    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m')
+    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
   \]
@@
     assert r1 1
-    halt
-2f:
+2f: halt
+
 \end{verbatim}
 
+\begin{lemma}
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+2) \\
+    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-2,\mathtt{2f},\mathtt{f2}) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
+    m & \defeq \hs_{f1} \uplus 
+               \hs_\flag \uplus                
+               \ms_{\var{link}} \uplus 
+               \hs_\adv \uplus 
+               m_{\malloc} \uplus 
+               \ms_{\var{stk}}
+               \ms_{\var{frame}} 
+  \end{align*}
+where 
+\begin{align*}
+  &\dom(\hs_{f2}) = [\mathtt{f2}-2,\mathtt{2f}] \\
+  &\dom(\hs_\flag) = [\flag,\flag] \\
+  &\dom(\ms_\link) = [\link,\link+1]\\
+  &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
+\end{align*}
+and
+\begin{itemize}
+\item The first address of $\hs_{f2}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f2}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f2}$ contains the code of $f2$.
+\item $\ms_\flag = [flag \mapsto 0]$
+\item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+\item $\ms_\stk(a) = 0$ for all $a \in [\start_\stk, \addrend_\stk]$
+\item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
+\end{itemize}
+if 
+  \[
+    (\reg_0\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+  \]
+then
+\[
+  m'(\flag) = 0
+\]  
+\end{lemma}
+
+ML-like program:
 \begin{verbatim}
@@
 Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
-
-Correctness lemma not assuming stack
-\begin{lemma}
-  For all $n$,
-  let
-  \begin{align*}
-    c_\adv & = ((\entry,\local),\start_\adv,\addrend_\adv,\start_\adv) \\
-    c_f & = ((\exec,\local),\start_f,\addrend_f,\start_f)
-  \end{align*}
-  if
-  \[
-    (\reg_0\update{\pcreg}{c_f}\update{r_1}{c_\adv},\hs_\adv \uplus \hs_f \uplus m_\malloc \uplus \hs_\flag \uplus \hs_{\var{frame}}) \step[n] (\halted, h')
-  \]
-  where
-\lau{TODO: finish this}
-\end{lemma}
 
@@
 We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
-\lars{don't you want to allocate the location \texttt{x} first and then
-  return a function taking \texttt{adv} as an argument ??}
-\lau{I am not sure I get what you have in mind, could you perhaps write the program you had in mind?}
 

### 2016-11-23 local-capability-machine.tex
Added clear macros

@@
 \end{align*}
-Further, assume an inverse function, $\decodePermPair$, that decodes permissions
+Further, assume an inverse function, $\decodePermPair{}$, that decodes permissions
 \[
-  \decodePermPair : \ints \fun (\Perms \times \Globals)
+  \decodePermPair{} : \ints \fun (\Perms \times \Globals)
 \]
@@
 \end{align*}
-It is assumed that address 0 is used for I/O, so whatever resides in location 0 after an execution can be seen as a result.
-As we will see in the semantics, it is assumed that the initial process starts with a readwrite capability for this location.
+
+A number of functions and predicates used in the definition of $\sem{-}$ (defined later). Notice all of them are total.
 \begin{align*}
@@
 If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
+\item[cclear $r$] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
+In a real system it would probably be implemented as a vector operation which allows modification of continous segments of memory rather fast.}
+\item[rclear $\bar{r}$] Moves 0 to all the registers in the list $\bar{r}$.
 \end{description}
@@
 If any callbacks are provided, then they need to be global.
+\item In a real setting due to a limited number of registers, some of the arguments might be spilled to the stack. It would be possible to do something similar here, but to keep
+ matters simple, we opt not to do so.
 \end{itemize}

### 2016-11-23 local-capability-machine.tex
Minor text fixes

@@
 \subsubsection{Simple use of local capabilities}
+The following programs only use local capabilities when they create protected return pointers. The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
+
 ML-like program we had before:
@@
 \end{verbatim}
+\lau{The above program is obsolete (and the other ML-like programs), but I will not remove it until we have gotten used to the assembly programs.}
+
 Assembly program not using stack. Assume that $r_l \not\in \{\pcreg,r_0 \}$ is a register.
@@
 1f: halt
-
 \end{verbatim}
+
 \begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
@@
 
-Assembly program using the stack:
+Assembly program using the stack. This program assumes a $r_\stk \not\in \{\pcreg,r_0\}$ register that contains a stack capability (a local $rwlx$-capability):
 \begin{verbatim}
@@
 2f: halt
-
 \end{verbatim}

### 2016-11-23 local-capability-machine.tex
Added an empty proof
Updated Malloc specification so the footprint is a memory rather than a memory segment

@@
   \begin{align*}
-    &\forall \Phi \in \ExecConfs \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
-    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
+    &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
+    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
     &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
     &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
-%    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\ % This
-    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
-    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
+    &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
+    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
     &\qquad \quad \memreg(r_1) = \size \land \\
@@
     &\qquad \quad \Rightarrow \\
-    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
+    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \HeapSegments\ldotp\\
     &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
     &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
-    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot'\uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+    &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
     &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
-    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
+    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
     &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
@@
 \end{specification}
-\lau{Only configurations with a memory (not memory segments) are executable.
-Therefore either $\hsfoot$ or $\hsframe$ need to be a memory and not a memory segment (it should probably be the footprint?).}
 
@@
 \begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
+\label{lem:correctness-f1}
   For all $n \in \nats$
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-f1}]
+
+\end{proof}
 

### 2016-11-23 local-capability-machine.tex
Changed the malloc specification so it leaves unused registers unchanged.

@@
     &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
-    &\qquad \qquad \qquad \size = e'-b' \\
-    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
-    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
+    &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+    &\qquad \qquad \qquad \size = e'-b' \land\\
   \end{align*}
@@
 In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
-\dominique{how is the return capability constructed: not on the stack, I presume, so this means mallocing a block of memory for the activation record?}
+\dominique{how is the return capability constructed: not on the stack, I presume, so  means mallocing a block of memory for the activation record?}
 \lau{See the new addition below - when we don't have a stack, then we have no need to ``save'' local capabilities which means that we can have the return code as part of the program.

### 2016-11-23 local-capability-machine.tex
More adjustments to the malloc specification

@@
   $c_\malloc$ satisfies the specification for malloc iff
-  \begin{align*}
+\[  
+  \left\{
+  \begin{aligned}
     &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
@@
     &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
-    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
+    &\qquad \quad \forall W \futurewk [i \mapsto \iota_{\malloc,0}] \land \\
     &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
@@
     &\qquad \quad \Rightarrow \\
-    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \HeapSegments\ldotp\\
+    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \Heaps\ldotp \exists \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
     &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
@@
     &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
+    &\qquad \qquad \qquad W'(i) = \iota_{\malloc}' \land W' \futurewk W \land \\
     &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
@@
     &\qquad \qquad \qquad \size = e'-b' \land\\
-  \end{align*}
-
-  Additionally, we require that for every $W' \future [i\mapsto \iota_{\malloc,0}]$, $\heapSat[\hs]{n}{W'}$ implies that $\dom(\hs)$ contains $[n',+\infty)$ for some $n'$.
+  \end{aligned}\right.
+\]
+where $\iota_{\malloc,0}$ is the region that initially governs malloc 
 \end{specification}
+In the specfication above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
 
@@
 \item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
+
 \item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison).

### 2016-11-23 local-capability-machine.tex
Changed malloc specification and added a bit of a proof

@@
     &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
-    &\qquad \quad \forall W \futurewk [i \mapsto \iota_{\malloc,0}] \land \\
+    &\qquad \quad \forall \iota_\malloc \futurewk \iota_{\malloc,0} \land \\
     &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
-    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
+    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
     &\qquad \quad \memreg(r_1) = \size \land \\
@@
     &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-    &\qquad \qquad \qquad W'(i) = \iota_{\malloc}' \land W' \futurewk W \land \\
-    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
+    &\qquad \qquad \qquad  \iota_{\malloc}' \futurewk \iota_{\malloc}\\
+    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{[i \mapsto \iota_{\malloc}']} \land \\
     &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
@@
   &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \text{ for all } n
 \end{align*}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f1}]
+Let $n$ be given and assume the premisses in the lemma.  Consider the following part of the execution:
+  \[
+    (\reg_0\update{\pcreg}{c_{f1}},m) \step[i] (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}'}\update{r_1}{1},m)
+  \]
+Where $c_{f1}'$ is the return address. Use the malloc specification with
+\begin{align*}
+  W &= [0 \mapsto \iota_{\malloc,0}] \\
+  m_{\var{footprint}} & = m_\malloc \\
+  size & = 1
+\end{align*}
+to get 
+\[
+(\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}'}\update{r_1}{1},m) \step[j] (\reg_0\update{\pcreg}{c_{f1}'}\update{r_0}{c_{f1}'}\update{r_1}{c_l},m')
+\]
+where for some $W' \futurewk W$
+\begin{align*}
+  m' & = \hs_{f1} \uplus 
+         \hs_\flag \uplus                
+         \ms_{\var{link}} \uplus 
+         \hs_\adv \uplus 
+         \ms_{\var{alloc}} \uplus
+         m_{\malloc}' \uplus 
+         \hs_{\var{frame}} 
+\end{align*}
 
@@
     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \text{ for all } n
   \end{align*}

### 2016-11-24 local-capability-machine.tex
Small change in malloc spec and semantics

@@
                                                                 \text{if $\memreg(\pcreg) = \stdcap$}\\
-                                                                \quad\text{and $\start \leq \addr < \addrend$}\\
+                                                                \quad\text{and $\start \leq \addr \leq \addrend$}\\
                                                                 \quad\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$ }
@@
     &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
-    &\qquad \quad \forall \iota_\malloc \futurewk \iota_{\malloc,0} \land \\
+    &\qquad \quad \forall W \futurewk [i \mapsto \iota_{\malloc,0}] \land \\
     &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
-    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
+    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
     &\qquad \quad \memreg(r_1) = \size \land \\
@@
     &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-    &\qquad \qquad \qquad  \iota_{\malloc}' \futurewk \iota_{\malloc}\\
-    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{[i \mapsto \iota_{\malloc}']} \land \\
+    &\qquad \qquad \qquad W'= W \update{i}{\iota_{\malloc}'} \land W' \futurewk W \land \\
+    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
     &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
     &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-    &\qquad \qquad \qquad \size = e'-b' \land\\
+    &\qquad \qquad \qquad \size - 1 = e'-b' 
   \end{aligned}\right.
@@
 In the specfication above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
+\lau{should it be a public or private future world in the premise?}
 
@@
   m_{\var{footprint}} & = m_\malloc \\
-  size & = 1
+  \memreg(r_1) & = \size = 1
 \end{align*}
@@
 \]
-where for some $W' \futurewk W$
-\begin{align*}
-  m' & = \hs_{f1} \uplus 
+for some $j$ where for some $[0 \mapsto \iota_{\malloc}] \futurewk [0 \mapsto_{\malloc,0}]$
+\begin{enumerate}
+\item $m' = \hs_{f1} \uplus 
          \hs_\flag \uplus                
@@
          m_{\malloc}' \uplus 
-         \hs_{\var{frame}} 
-\end{align*}
+         \hs_{\var{frame}} $
+\item $\heapSat[m_{\malloc}']{n-j}{[0 \mapsto \iota_{\malloc}]}$
+\item $\dom[\hs_{\var{alloc}}] = [l,l]$
+\item $c_l = ((\rwx,\glob),l,l,l)$
+\end{enumerate}
 

### 2016-11-24 local-capability-machine.tex
Updated malloc specification

@@
 \documentclass[a4paper]{article}
-%\usepackage[a4paper]{geometry}
+\usepackage[a4paper]{geometry}
 \usepackage[utf8]{inputenc}
@@
 
+\usepackage{enumitem}
 \usepackage{tikz}
@@
 \[  
-  \left\{
   \begin{aligned}
-    &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
-    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
-    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
-    &\qquad \quad \forall W \futurewk [i \mapsto \iota_{\malloc,0}] \land \\
-    &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
-    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
+    & \exists \iota_{\malloc,0} \ldotp \\
+    & \quad \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
+    &\qquad \forall i, n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \\
+    &\qquad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
+    &\qquad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land\\
+    &\qquad \quad \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
     &\qquad \quad \memreg(r_1) = \size \land \\
@@
     &\qquad \quad \Rightarrow \\
-    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \Heaps\ldotp \exists \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
-    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
-    &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-    &\qquad \qquad \qquad W'= W \update{i}{\iota_{\malloc}'} \land W' \futurewk W \land \\
-    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
+    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
+    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+    &\qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
+    &\qquad \qquad \qquad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+    &\qquad \qquad \qquad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
+    &\qquad \qquad \qquad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
     &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
@@
     &\qquad \qquad \qquad \size - 1 = e'-b' 
-  \end{aligned}\right.
+  \end{aligned}
 \]
 where $\iota_{\malloc,0}$ is the region that initially governs malloc 
+\[
+\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i, W(i)=\iota' \ldotp \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))
+\]
+and
+\[
+  \iota_{\malloc,0}.v = \perm
+\]
+
 \end{specification}
@@
     \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
-    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
+    \item Adjust a copy of the current pc to point to the return address and save it to the activation record.
     \item Write the activation code to the activation record.
@@
 \]
-for some $j$ where for some $[0 \mapsto \iota_{\malloc}] \futurewk [0 \mapsto_{\malloc,0}]$
+for some $j$ where for some $[0 \mapsto \iota_{\malloc}] \futurewk [0 \mapsto \iota_{\malloc,0}]$
 \begin{enumerate}
@@
          \hs_\adv \uplus 
-         \ms_{\var{alloc}} \uplus
+         \ms_l \uplus
          m_{\malloc}' \uplus 
          \hs_{\var{frame}} $
-\item $\heapSat[m_{\malloc}']{n-j}{[0 \mapsto \iota_{\malloc}]}$
-\item $\dom[\hs_{\var{alloc}}] = [l,l]$
-\item $c_l = ((\rwx,\glob),l,l,l)$
+\item $\heapSat[m_{\malloc}']{n-j}{[0 \mapsto \iota_{\malloc}]}$ \label{f1:mallocsat}
+\item $\dom(\hs_l) = [l,l]$
+\item $c_l = ((\rwx,\glob),l,l,l)$ \label{test}
+\item $\ms_l(l) = 0$
 \end{enumerate}
+Continue the execution to the next malloc hidden in \texttt{call}.
+\[
+  (\reg_0\update{\pcreg}{c_{f1}'}\update{r_0}{c_{f1}'}\update{r_1}{c_l},m')
+  \step[k]
+  (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}''}\update{r_1}{\dots}\update{r_l}{c_l},m'')
+\]
+where
+\begin{enumerate}[resume]
+\item $m'' = m'[l\mapsto 1]$
+\end{enumerate}
+Use the malloc specification notice:
+\begin{itemize}
+\item $\cdots$ represents the needed size for the activation record. 
+\item \ref{f1:mallocsat}.\ and (downwards closure) gives us the needed memory segment satisfaction.
+\item $m_{\var{footprint}} =  m_\malloc' $
+\end{itemize}
+Get:
+\[
+  (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}''}\update{r_1}{\dots}\update{r_l}{c_l},m'')
+\step[j']
+  (\reg_0\update{\pcreg}{c_{f1}''}\update{r_0}{c_{f1}''}\update{r_1}{c_{\var{ar}}}\update{r_l}{c_l},m^{(3)})
+\]
+for some $j'$ where for some $[0 \mapsto \iota_{\malloc}'] \futurewk [0 \mapsto \iota_{\malloc}]$
+\begin{enumerate}[resume]
+\item $m'' = \hs_{f1} \uplus 
+         \hs_\flag \uplus                
+         \ms_{\var{link}} \uplus 
+         \hs_\adv \uplus 
+         \ms_{l} \uplus
+         \ms_{\var{ar}} \uplus
+         m_{\malloc}'' \uplus 
+         \hs_{\var{frame}} $
+\item $\heapSat[m_{\malloc}'']{n-j}{[0 \mapsto \iota_{\malloc}']}$ \label{f1:mallocsat}
+\item $\dom(\hs_{\var{ar}}) = [b,e]$
+\item $c_l = ((\rwx,\glob),b,e,b)$ \label{test}
+\item $\forall a \in [b,e] \ldotp \ms_{\var{ar}}(a) = 0$
+\end{enumerate}
+Contniue execution until just after the jump to $\var{adv}$.
+\[
+  (\reg_0\update{\pcreg}{c_{f1}''}\update{r_0}{c_{f1}''}\update{r_1}{c_{\var{ar}}}\update{r_l}{c_l},m^{(3)})
+\step[k']
+  (\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_0}{c_{ar}'}\update{r_l}{c_l},m^{(3)})
+\]
+for some $k'$ wher
+\begin{itemize}
+\item $\ms_{\var{ar}}$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$, and activation code.
+\item $c_{ar}' = ((\entry,\local)b,e,b+2)$ where $b+2$ is the first address of the activation code.
+\end{itemize}
 

### 2016-11-25 local-capability-machine.tex
Changes to malloc for readability
f1 proof

@@
 \usepackage{enumitem}
+\newlist{enumproof}{enumerate}{10}
+\setlist[enumproof]{label*=\arabic*.}
+
 \usepackage{tikz}
@@
 \newcommand{\flag}{\var{flag}}
+\newcommand{\nwl}{\var{nwl}}
+\newcommand{\pwl}{\var{pwl}}
 
@@
 \[
-\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i, W(i)=\iota' \ldotp \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))
+\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))
 \]
@@
 \[
-  \iota_{\malloc,0}.v = \perm
+  \iota_{\malloc,0}.v = \perma
 \]
@@
 \end{specification}
-In the specfication above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
+In the specification above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
 \lau{should it be a public or private future world in the premise?}
@@
 If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
-\item[cclear $r$] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
-In a real system it would probably be implemented as a vector operation which allows modification of continous segments of memory rather fast.}
-\item[rclear $\bar{r}$] Moves 0 to all the registers in the list $\bar{r}$.
+\item[\texttt{cclear $r$}] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
+In a real system it would probably be implemented as a vector operation which allows modification of continuous segments of memory rather fast.}
+\item[\texttt{rclear $\bar{r}$}] Moves 0 to all the registers in the list $\bar{r}$.
 \end{description}
@@
 
-The stack will reside entirely in memory. There is no seperation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
-
-Eventhough the memory is infinite, we will only use a finite part for the stack.
+The stack will reside entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
+
+Even though the memory is infinite, we will only use a finite part for the stack.
 If we have allocated too little memory for the stack, and we try to push something anyway, then the execution will fail.
-As we consider failing admissable, we are okay with this.
+As we consider failing admissible, we are okay with this.
 
@@
   \item [Push stack pointer]
-    Push the full stackpointer to the stack.
+    Push the full stack pointer to the stack.
   \item [Create protected return pointer]
@@
 In this example, the call \texttt{scall $f([r_{\var{args},1},\dots,r_{\var{args},n}],[r_0,r_{\var{priv},1},\dots,r_{\var{priv},m}])$.
-In this example the two lists of registers are disjoined eventhough that does not have to be the case.}}
+In this example the two lists of registers are disjoint even though that does not have to be the case.}}
 \end{figure}
@@
                \hs_\adv \uplus 
-               m_{\malloc} \uplus 
+               \ms_{\malloc} \uplus 
                \hs_{\var{frame}} 
@@
 \end{lemma}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f1}]
-Let $n$ be given and assume the premisses in the lemma.  Consider the following part of the execution:
+Let $n$ be given and assume the premises in the lemma.  Consider the following part of the execution:
   \[
@@
 \begin{align*}
-  W &= [0 \mapsto \iota_{\malloc,0}] \\
-  m_{\var{footprint}} & = m_\malloc \\
+  \iota_\malloc &= \iota_{\malloc,0} \\
+  \ms_{\var{footprint}} & = \ms_\malloc \\
   \memreg(r_1) & = \size = 1
@@
 \]
-for some $j$ where for some $[0 \mapsto \iota_{\malloc}] \futurewk [0 \mapsto \iota_{\malloc,0}]$
+for some $j$ where for some $\iota_{\malloc}' \futurewk \iota_{\malloc,0}$
 \begin{enumerate}
@@
          \ms_l \uplus
-         m_{\malloc}' \uplus 
+         \ms_{\malloc}' \uplus 
          \hs_{\var{frame}} $
-\item $\heapSat[m_{\malloc}']{n-j}{[0 \mapsto \iota_{\malloc}]}$ \label{f1:mallocsat}
+\item $\heapSat[\ms_{\malloc}']{n-j}{[0 \mapsto \iota_{\malloc}]}$ \label{f1:mallocsat}
 \item $\dom(\hs_l) = [l,l]$
@@
 \item \ref{f1:mallocsat}.\ and (downwards closure) gives us the needed memory segment satisfaction.
-\item $m_{\var{footprint}} =  m_\malloc' $
+\item $\ms_{\var{footprint}} =  \ms_\malloc' $
 \end{itemize}
@@
          \ms_{\var{ar}} \uplus
-         m_{\malloc}'' \uplus 
+         \ms_{\malloc}'' \uplus 
          \hs_{\var{frame}} $
-\item $\heapSat[m_{\malloc}'']{n-j}{[0 \mapsto \iota_{\malloc}']}$ \label{f1:mallocsat}
+\item $\heapSat[\ms_{\malloc}'']{n-j-j'}{[0 \mapsto \iota_{\malloc}']}$ \label{f1:mallocsat}
 \item $\dom(\hs_{\var{ar}}) = [b,e]$
@@
 \end{enumerate}
-Contniue execution until just after the jump to $\var{adv}$.
+Continue execution until just after the jump to $\var{adv}$.
 \[
@@
 \step[k']
-  (\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_0}{c_{ar}'}\update{r_l}{c_l},m^{(3)})
+  (\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_1}{c_{\var{adv}}}\update{r_0}{c_{ar}'},m^{(3)})
 \]
-for some $k'$ wher
+for some $k'$ where
 \begin{itemize}
-\item $\ms_{\var{ar}}$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$, and activation code.
-\item $c_{ar}' = ((\entry,\local)b,e,b+2)$ where $b+2$ is the first address of the activation code.
+\item $m^{(3)} = \hs_{f1} \uplus 
+         \hs_\flag \uplus                
+         \ms_{\var{link}} \uplus 
+         \hs_\adv \uplus 
+         \ms_{l} \uplus
+         \ms_{\var{ar}}' \uplus
+         \ms_{\malloc}'' \uplus 
+         \hs_{\var{frame}} $
+\item $\ms_{\var{ar}}'$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$ (the return address in f1), and activation code.
+\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+2)$ where $b+2$ is the first address of the activation code.
 \end{itemize}
-
+Define
+\begin{itemize}
+\item $W = [0 \mapsto \iota_\malloc']
+           [1 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv}(\perma)]
+           [2 \mapsto \iota_{f1}]
+           [3 \mapsto \iota_{\var{ar}}]
+           [4 \mapsto \iota_l]
+           [5 \mapsto \iota_\flag]
+           [5 \mapsto \iota_\link]$
+\end{itemize}
+where
+\begin{itemize}
+\item $\iota_{f1}$, $\iota_{\var{ar}}$, $\iota_l$, $\iota_\flag$, and $\iota_\link$ are permanent islands that only accepts respectively $\ms_{f1}$, $\ms_{\var{ar}}$, $\ms_l$, $\ms_\flag$, and $\ms_\link$.
+\end{itemize}
+define
+\begin{enumerate}
+\item $\ms = \hs_{f1} \uplus 
+         \hs_\flag \uplus                
+         \ms_{\var{link}} \uplus 
+         \hs_\adv \uplus 
+         \ms_{l} \uplus
+         \ms_{\var{ar}}' \uplus
+         \ms_{\malloc}'' $ \label{f1:ms-disj-union}
+\end{enumerate}
+Use the FTLR on $\updatePcPerm{c_{\var{adv}}}$ using world $W$, so show
+\begin{itemize}
+\item $\npair{(\start_{\adv},\addrend_{\adv})} \in \readCond{}(\glob)(W)$
+  \begin{itemize}
+  \item Show: $\mapsto \iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.
+  \end{itemize}
+\end{itemize}
+Have
+\begin{enumerate}[resume]
+\item $\npair{\updatePcPerm{c_{\var{adv}}}} \in \stder(W)$
+\end{enumerate}
+Let $n' = n - j - j'$ and show
+\begin{enumproof}[resume]
+\item $n' \leq n$
+  \begin{enumproof}
+    \item trivial
+  \end{enumproof}
+\item $\heapSat[\ms]{n'}{W}$
+  \begin{enumproof}
+    \item Split the memory into the disjoint unions of \ref{f1:m3-disj-union} and show:
+      \begin{enumproof}
+      \item case: $\npair[n']{\ms_\adv} \in \iota_\malloc'.H (\iota_\malloc'.s) (W)$
+        \begin{enumproof}
+          \item Use $\heapSat[\ms_\malloc]{n'}{[0 \mapsto \iota_\malloc']}$ with malloc specification context independence property.
+        \end{enumproof}
+      \item case: $\npair[n']{\ms_\adv} \in H^\nwl_{\start_\adv,\addrend_\adv} (\start_\adv,\addrend_\adv) W$
+        \begin{enumproof}
+          \item Show $\forall a \in [\start_\adv,\addrend_\adv] \ldotp \npair[n'-1]{\ms(a)} \in \stdvr(\revokeTemp{W})$
+          \item Notice: $W = \revokeTemp(W)$
+        \end{enumproof}
+        \begin{enumproof}
+          \item $a \neq \start_\adv$ : trivial, contains instruction only.
+          \item $a = \start_\adv$: show $((\readonly,\glob),\link,\link+1,\link) \in \stdvr(W)$\\
+           SFTS $\iota_{\link} \nsubeq[n'-1] \iota^\pwl_{\link,\link+1} (\perma)$\\
+           $\iota_{\link}$ accepts only $\ms_\link$, so show:
+            \begin{enumproof}
+            \item $\npair[n'-2]{c_\malloc} \in \stdvr(W)$: We assume this is indeed the case.
+            \item $\npair[n'-2]{c_\adv} \in \stdvr(W)$: for $n'' < n'-2$ and $W' \futurestr W$ show: \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stdvr(W')$. Use FTLR. \label{f1:adv}
+            \end{enumproof}
+          \end{enumproof}
+      \item The remaining islands are designed to accept their designated islands.
+    \end{enumproof}
+  \end{enumproof}
+\item $\npair[n']{\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_1}{c_{\var{adv}}}\update{r_0}{c_{ar}'}} \in \stdrr(W)$
+  \begin{enumproof}
+    \item case: $\npair[n']{c_{\var{adv}}} \in \stdvr(W)$
+      \begin{enumproof}
+        \item Similar to \ref{f1:adv}
+      \end{enumproof}
+    \item case: $\npair[n']{c_{\var{ar}}'} \in \stdvr(W)$.
+      \begin{enumproof}
+        \item Let $n'' < n'$ and $W' \futurewk W$ be given and show $\npair[n'']{\updatePcPerm{c_{\var{ar}}'}} \in \stdvr(W')$\\
+          Let $n^{(3)} \leq n''$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n^{(3)}]{\reg}$ be given\\
+          Show: $\npair[n^{(3)}]{(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms)}$\\
+          Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k''$, $m'$ and $\ms_{\var{frame}}$.
+Due to $\heapSat[\ms']{n''}{W'}$ $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
+\\
+          The execution loads $c_l$ to $r_l$ and jumps to $c_{f1}^{(3)}$ (the point just before the assertion).
+As $\ms_l = 1$, the assetion is successful and the execution halts.
+In other words, there were no changes to the memory.\\
+          Use $W'$, $\ms_r = \emptyset$, and $\ms'$ to get the desired result, i.e., $m' = \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n''-k''}{W'}$ (using downwardsclosure of memory satisfaction).
+      \end{enumproof}
+    \item case: $\npair[n']{0} \in \stdvr(W)$
+      \begin{enumproof}
+        \item Trivial
+      \end{enumproof}
+  \end{enumproof}
+\end{enumproof}
+Get
+\[
+\npair[n']{(\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_1}{c_{\var{adv}}}\update{r_0}{c_{ar}'},\ms)} \in \observations(W)
+\]
+By initial assumption of the lemma, the execution halts. Use $\ms_{\var{frame}}$, $m'$ and the number of steps it takes to halt to get:
+$W' \futurestr W$, $\ms_r$ and $\ms'$ s.t.
+$m' = \ms_r \uplus \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n' - \dots}{W'}$.
+As $\iota_\flag$ is a permanent region, we know it is still in $W'$, so $m'(\flag) = 0$.
 \end{proof}
 
-Assembly program using the stack. This program assumes a $r_\stk \not\in \{\pcreg,r_0\}$ register that contains a stack capability (a local $rwlx$-capability):
+Assembly program using the stack. This program assumes a $r_\stk \not\in \{\pcreg,r_0\}$ register that contains a stack capability (a local $\rwlx$-capability):
 \begin{verbatim}
@@
         \hs_\adv \uplus 
-        m_{\malloc} \uplus 
+        \ms_{\malloc} \uplus 
         \ms_{\var{stk}}
@@
 \\If it looks like a stack, works like a stack, and quacks like a stack, then it is probably a stack.
-The first requirement makes sure that what was passed as the stack can be used as a stack eventhough it might not be a stack capability that follows the conventional stack diciplin.
+The first requirement makes sure that what was passed as the stack can be used as a stack even though it might not be a stack capability that follows the conventional stack discipline.
 If the capability is actually for part of the callers stack, then we will just overwrite their part of the stack.
@@
 This is the case as the \emph{only} place they could have stored the stack capability is on the stack, and the only way they can restore this is through one of the local stack capabilities.
-In other words, this requirement makes sure that the callback cabaility is not derived from the stack pointer (which would allow them to get a stack capability for our part of the stack).\\ The last requirement is to sanitize the stack.
+In other words, this requirement makes sure that the callback capability is not derived from the stack pointer (which would allow them to get a stack capability for our part of the stack).\\ The last requirement is to sanitise the stack.
 This makes sure that the caller does not try to sneak a capability to the callback simply by leaving it somewhere on the stack and hoping that we won't notice and pass it on.
@@
     (\reg_0\update{\pcreg}{c_{\adv}}
-          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus m_f) \step[n] (\halted,m')
+          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus \hs_{\malloc} \uplus \ms_f) \step[n] (\halted,m')
   \]
@@
                              \begin{aligned}[t]
-                               & \forall \heap_f, \heap', i \leq n \ldotp \\
-                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
+                               & \forall \ms_f, \heap', i \leq n \ldotp \\
+                               & \quad(\reg,\hs \uplus \ms_f) \step[i] (\halted,\heap')  \\
                                & \qquad \Rightarrow
@@
                                  & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
-                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \heap_f \land \\ 
+                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \ms_f \land \\ 
                                  & \quad \heapSat[\hs']{n-i}{W'} \}

### 2016-11-25 local-capability-machine.tex
Added comment in scall
Changed scall

@@
   \item Restrict stack capability
+  \item Clear the stack we release control over
   \item Clear unused registers
@@
   \item Upon return: Run restore code
-  \item Return address: Clear the stack segment we gave control over
+  \item Return address: 
   \item Restore ``private'' state
@@
     Make the stack capability only govern the unused part.
+  \item [Clear the stack we release control over]
+    Store 0 to all the memory cells the restricted stack pointer has authority over.
   \item [Clear unused registers]
@@
     \begin{description}
-    \item [Clear the stack segment we gave control over] Store 0 to all the cells the stack capability passed to the callee governed.
     \item [Restore ``private'' state] \forcenewline
@@
 \cline{2-2}
+   & $?$\\
+\cline{2-2}
 $c_{\var{stk}}'' \rightarrow$  & $c_{\var{stk}}'$ \\
@@
 \caption{ Stack and register-file after the $c_{\var{stk}}'$ has been limited to only give authority over the empty part of the stack (the new capability is $c_{\var{stk}}''$).
+The empty part of the stack has been cleared.
 $c_0'$ is made from $c_{\var{stk}}'$ by setting it to point to the restore code and restricting it to a local enter-capability.
@@
    &  \\
+   & $\vdots$ \\
+\cline{2-2}
+   & ?\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
    & $\vdots$\\
@@
 \end{tabular}
-\caption{ Stack and register-register file after the clean up code has been run.
-The untrusted part of the stack has been cleared.
-The ``private'' words have been poped to their respective registers.
-The restore code has been poped of the stack.
-The stack is now in the same state as it was before the call.
-}
+\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been poped to their respective registers. The restore code has been poped of the stack. }
 \end{figure}

### 2016-11-28 local-capability-machine.tex
Updated f2 correctness lemma
small corrections

@@
 \newcommand{\pwl}{\var{pwl}}
+\newcommand{\sta}{\var{sta}}
+\newcommand{\cnst}{\var{cnst}}
 
@@
 \newcommand{\Heaps}{\plaindom{Mem}}
+\newcommand{\Mems}{\Heaps}
 \newcommand{\HeapSegments}{\plaindom{MemSegment}}
+\newcommand{\MemSegments}{\HeapSegments}
 \newcommand{\Confs}{\plaindom{Conf}}
@@
 \ExecConfs  &::= \Regs \times \Heaps \\
+\Globals & ::= \{\glob, \local \} \\
 \Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
-\Confs &::= \ExecConfs + \{\failed, \halted \times \ExecConfs\} \\
-\HeapSegments &::= \Addrs \parfun \Words\\
-\Globals & ::= \{\glob, \local \}
+\Confs &::= \ExecConfs + \{\failed \} + \halted \times \ExecConfs \\
+\HeapSegments &::= \Addrs \parfun \Words
 \end{align*}
 Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capability is local or global.
-Two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities.
+There are two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities.
 They are otherwise the same as their non-''permit write local'' counterparts.
 
-The $\glob$ and $\local$ are ordered as shown in Figure~\ref{fig:glob-hier}.
 \begin{figure}[!h]
@@
 \hline
-  $\readwrite$ & Read-write. Grants read and write permissions. Storage of local capabilities prohibited. \\
+  $\readwrite$ & Read-write. Grants read and write permission. Storage of local capabilities prohibited. \\
 \hline
-  $\readwritel$ & read-write, permit write local. Grants read and write permissions. Storage of local capabilities granted. \\
+  $\readwritel$ & Read-write, permit write local. Grants read and write permission. Storage of local capabilities possible. \\
 \hline
@@
 \hline
-  $\entry$ & Entry permission. This permission grants no access, but when jumped to, it Will turn into an $\exec$ permission.\\
+  $\entry$ & Enter permission. This permission grants no access, but when jumped to, it will turn into an $\exec$ permission.\\
 \hline
-  $\rwx$ & Read, write, execute permission. Grants read, write, and execute permissions. Storage of local capabilities prohibited. \\
+  $\rwx$ & Read-write-execute permission. Grants read, write, and execute permissions. Storage of local capabilities prohibited. \\
 \hline
-  $\rwlx$ & Read, write, execute permission with permit write local. Grants read, write, and execute permissions. Storage of local capabilities possible.
+  $\rwlx$ & Read-write-execute, permit write local. Grants read, write, and execute permissions. Storage of local capabilities possible.
 \end{tabular}
@@
 \Phi    &\in& \ExecConfs \\
-\memheap&\in& \Heaps \\
+m, \memheap&\in& \Heaps \\
 \memreg &\in& \Regs \\
@@
 n       &\in& \ints\\
+\ms     &\in& \MemSegments
 \end{array}
 \]
-Further definitions:
+Words and instructions:
 \[
@@
 \subsection*{Semantics}
-Assume a $\decode$ function that decodes integer to instructions:
+Assume a $\decode$ function that decodes words to instructions:
 \begin{align*}
@@
                                                    \quad\text{and in either case $n \in \ints$}\\
-                                                   \quad\text{and $\var{newPermPair} = \decodePermPair(n)$}\\
-                                                   \quad\text{and $\var{newPermPair} \sqsubseteq \permp$}\\
-                                                   \quad\text{and $c = (\var{newPermPair},\start,\addrend,\addr)$}
+                                                   \quad\text{and $\decodePermPair{n}\sqsubseteq \permp$}\\
+                                                   \quad\text{and $c = (\decodePermPair{n},\start,\addrend,\addr)$}
                                                  \end{array}\\
@@
 \end{align*}
+\lau{Do we need other arithmetic operators than plus?}
 \begin{align*}
@@
                                               \begin{array}[t]{l}
-                                                \text{if $\memreg{r} = \stdcap[\permp]$} \\
+                                                \text{if $\memreg(r) = \stdcap[\permp]$} \\
                                                 \quad\text{and for $i \in \{1,2\}$}\\
@@
                                                 \quad\text{and $\start \leq n_1 \leq n_2 \leq \addrend$}\\
-                                                \quad\text{and $c = (\permp,n_2,n_2,\addr)$}
+                                                \quad\text{and $c = (\permp,n_1,n_2,\addr)$}
                                               \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg{r_2} = \stdcap$}
+                              \text{if $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg{r_2} = \stdcap$}
+                              \text{if $\memreg(r_2) = ((\_,\_),\start,\_,\_)$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg{r_2} = \stdcap$}
+                              \text{if $\memreg(r_2) = ((\_,\_),\_,\addrend,\_)$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg{r_2} = \stdcap$}
+                              \text{if $\memreg(r_2) = ((\perm,\_),\_,\_,\_)$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg{r_2} = \stdcap$}
+                              \text{if $\memreg(r_2) = ((\_,\gl),\_,\_,\_)$}
                             \end{array} \\
@@
                           \begin{cases}
-                            \stdUpdatePc{\updateReg{r_1}{1}} & \text{if $\memreg{\rv} = \stdcap$} \\
+                            \stdUpdatePc{\updateReg{r_1}{1}} & \text{if $\memreg(\rv) \in \Caps$ } \\
                             \stdUpdatePc{\updateReg{r_1}{0}} & \text{otherwise} 
@@
 
-After adding local capabilities, most of the definitions have changed slightly to take into the account new form the capabilities take.
-The only two instructions that have undergone enough change to deserve to be mentioned are $\store{}$ and $\restrict{}{}{}$.
-
-$\store{}$ has the same conditions as before, but if what it tries to write is a capability, then it needs to be a write access that permits write local.
-
-$\restrict{}{}{}$ looks like before, but now it uses a ``permission pair'' which is the access permission and whether it is local or not.
-Whether a $\restrict{}{}{}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above.
-
-Define the following simple macros:
-
+After adding local capabilities, most of the definitions have changed slightly to take into account the new form the capabilities take.
+The only two instructions that have undergone enough change to deserve to be mentioned are $\mathtt{store}$ and $\mathtt{restrict}$.
+
+$\mathtt{store}$ has the same conditions as before, but if what we try to write is a local capability, then it needs to be a write access that permits write local.
+
+$\mathtt{restrict}$ looks like before, but now it uses a ``permission pair'' which is the access permission and the locality.
+Whether a $\mathtt{restrict}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above (it is the point-wise ordering of the pair).
+
+Define the following macros: $\mathtt{restrict}$, $\mathtt{subseg}$, and $\mathtt{lea}$ that does not overwrite the source register. A $\mathtt{store}$ that allows integers to be stored directly.
 \begin{align*}
@@
   \begin{aligned}
-    & \exists \iota_{\malloc,0} \ldotp \\
-    & \quad \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
+    &\exists \iota_{\malloc,0} \ldotp \\
+    &\quad (\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))) \; \land \\
+    &\quad \iota_{\malloc,0}.v = \perma \; \land \\
+    &\quad (\forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
     &\qquad \forall i, n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \\
@@
     &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+    &\qquad \qquad \quad \exists j \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
     &\qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
@@
     &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-    &\qquad \qquad \qquad \size - 1 = e'-b' 
+    &\qquad \qquad \qquad \size - 1 = e'-b' )
   \end{aligned}
 \]
-where $\iota_{\malloc,0}$ is the region that initially governs malloc 
-\[
-\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))
-\]
-and
-\[
-  \iota_{\malloc,0}.v = \perma
-\]
-
 \end{specification}
 In the specification above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
-\lau{should it be a public or private future world in the premise?}
 
@@
 In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
-\dominique{how is the return capability constructed: not on the stack, I presume, so  means mallocing a block of memory for the activation record?}
-\lau{See the new addition below - when we don't have a stack, then we have no need to ``save'' local capabilities which means that we can have the return code as part of the program.
-The only thing we have to do is to have the capability point to the return address and make the capability into an entry capability.
-}
-\lau{On further thought, conveniently the examples did not seem to need an activation record, but if we want \texttt{call} to work more generally speaking, then it needs to setup an activation record.
-I have changed the description of \texttt{call}, so it now uses an activation record.
-}
 \dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
@@
 In order to make capabilities to trusted code (and possibly untrusted code) available, we assume that some sort of linker has made these available.
-This is done in the following way: For every function, the first memory cell the capability for that function governs contains a capability for the list in question.
-Each function name in a program corresponds to an offset in the table, e.g., \texttt{malloc}.
-When a name is used in a program, it indicates what entry to pick from the list.
-The list should always be accessible by taking a copy of the capability in the $\pcreg$-register and adjusting it to point to the first cell it governs.
-
-The capability link list can be shared between multiple functions as it is accessed through a read-only capability.
-
-\subsection{Flag list}
+This is done in the following way: For every function, the first memory cell the capability for that function governs contains a capability for the linking table.
+Each function name in a program corresponds to an offset in the table, e.g., \texttt{malloc} could be at offset 0.
+When a name is used in a program, it indicates what entry from the linking table to pick.
+The table should always be accessible by taking a copy of the capability in the $\pcreg$-register and adjusting it to point to the first cell it governs.
+
+The capability linking table can be shared between multiple functions that are linked to the same capabiblities as it is accessed through read-only capabilities.
+
+\subsection{Flag table}
 A function may use flags to signal failure.
-We use the convention that a flag list is available in the second memory cell of a functions code (so just after the linking table).
-The flag list is accessed through a read-write capability and initially it contains all zero.
-Like the linking list, each entry is associated with a name which may appear in the macros.
-
-The flag list should never be shared between distrusting parties.
-
+We use the convention that a flag table is available in the second memory cell of a functions code (so just after the linking table).
+The flag table is accessed through a read-write capability and initially it contains all zero.
+Like the linking table, each entry is associated with a name which may appear in the macros.
+
+The flag table should never be shared between distrusting parties.
+
+\subsection{Macro definitions}
 \begin{description}
-\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
+\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline
+$\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
   \begin{itemize}
   \item Set up activation record
-  \item Create local enter capability for activation
+  \item Create local enter capability for activation (protected return pointer)
   \item Clear unused registers
@@
 \begin{description}
-  \item [Set up activation record]
+  \item [Set up activation record]\forcenewline
     \begin{itemize}
@@
     \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
-    \item Adjust a copy of the current pc to point to the return address and save it to the activation record.
+    \item Adjust a copy of the current pc to point to the return address in code and save it to the activation record.
     \item Write the activation code to the activation record.
   \end{itemize}
-  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability.
+  \item [Create local enter capability for activation] Adjust the capability for the activation record to point to the beginning of the activation record and restrict it to a local enter-capability.
 Place this capability in $r_0$.
@@
 %    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
-    \item Load the return address to $\pcreg$
+    \item Load the return capability to $\pcreg$
     \end{itemize}
 \end{description}
-\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
+\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register $r$.
 
@@
 If they are equal, then execution continues.
-If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
+If they are unequal, then the assertion flag named $\var{flag}$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
 \item[\texttt{cclear $r$}] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
@@
 \begin{itemize}
-\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local.
+\item \texttt{call} will fail if we have local capabilities in one of the registers of the ``private'' register list as it relies on a capability returned by malloc which will not be permit-write-local.
 Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
@@
 \subsection{Stack}
-Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma.
+Some programs will assume access to a stack which will be in part indicated by the program macros but also in the correctness lemma.
 The stack is accessed through a local $\rwlx$-capability.
@@
 
-The stack will reside entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
+The stack resides entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
 
@@
 When not in the middle of a push or a pop, the stack capability points to the top word of the stack.
-For an empty stack, the stack capability points to the address just outside of the range of authority for the stack capability.
+For an empty stack, the stack capability points to the address just below of the range of authority for the stack capability.
+
+The stack grows upwards
 
 \begin{description}
-\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the stack capability by one and storing the word through said capability.\footnote{The stack grows upwards.}
-\item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
-\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
+\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the address of the stack capability by one and storing the word through the stack capability.
+\item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, and decrementing the address of the stack capability.
+\item[\texttt{scall} \forcenewline
+$f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
 This call assumes $r_{\var{stk}}$ contains a stack capability.
@@
   \begin{itemize}
-  \item Push restore code to the stack.
+  \item Push the restore code to the stack.
   \item Push ``private'' registers to the stack.
   \item Push return address capability 
-  \item Push stack pointer
+  \item Push stack capability
   \item Create protected return pointer
-  \item Restrict stack capability
-  \item Clear the stack we release control over
+  \item Restrict stack capability to unused part 
+  \item Clear the part of the stack we release control over
   \item Clear unused registers
   \item Jump
-  \item Upon return: Run restore code
-  \item Return address: 
-  \item Restore ``private'' state
+  \item Upon return: Run the on stack restore code
+  \item Return address in caller-code:  Restore ``private'' state
   \end{itemize}
@@
 \begin{description}
-  \item [Push restore code to the stack]
-    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
+  \item [Push the restore code to the stack]
+    Push the restore code to the stack (described later).
+This code needs to be on the stack to make sure the stack capability can be restored.
+We keep the restore code on the stack minimal.
+The caller code does the rest of the restoration.
   \item [Push ``private'' registers to the stack]
@@
     Push a capability for the return address (in the memory) to the stack.
-  \item [Push stack pointer]
-    Push the full stack pointer to the stack.
+  \item [Push stack capability]
+    Push the full stack capability to the stack.
   \item [Create protected return pointer]
@@
 Restrict it to a local enter-capability and put it in $r_0$.
-\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?}
-  \item [Restrict stack capability]
+\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?} \lau{It is also used when we want to call code for the first time, so calling it a protected return pointer is more of a conceptul thing.}
+  \item [Restrict stack capability to unused part]
     Make the stack capability only govern the unused part.
-  \item [Clear the stack we release control over]
+  \item [Clear the part of the stack we release control over]
     Store 0 to all the memory cells the restricted stack pointer has authority over.
@@
   \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
-  \item [Restore code]
-    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
-  \item [Return Address:] \forcenewline
-    \begin{description}
-    \item [Restore ``private'' state] \forcenewline
+  \item [Run the on stack restore code]
+    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter (the return address in caller-code) from the stack to $\pcreg$.
+  \item [Return address in caller-code: Restore ``private'' state] \forcenewline
       \begin{itemize}
@@
       \item Pop the restore code of the stack
-      \end{itemize}
-  \end{description}
+      \end{itemize}\
 \end{description}
@@
 It is also possible to use a combination of \texttt{scall} and \texttt{call}, but when \texttt{call} is used, then we have no way to store the stack, so we cannot use \texttt{scall} after that.
-\item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared upon return.
+\item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared before it is passed to an untrusted party.
 \item As a rule of thumb: If you receive a stack from an untrusted source, then you need to check that it is a local $\rwlx$-capability and clear it!
@@
    & \\
-   & $\vdots$ \\
-\cline{2-2}
-   & ?\\
-\cline{2-2}
-   & 0 \\
-\cline{2-2}
    & $\vdots$\\
 \cline{2-2}
-   & 0 \\
-\cline{2-2}
-$c_{\var{stk}} \rightarrow$   & local stack\\
+   & ? \\
+\cline{2-2}
+$ \rightarrow$  & $c_0$ \\
+\cline{2-2}
+   & $w_{p,1}$ \\
+\cline{2-2}
+   & $\vdots$ \\
+\cline{2-2}
+   & $w_{p,m}$ \\
+\cline{2-2}
+   & Restore code \\
+   & $\vdots$\\
+\cline{2-2}
+ $c_\stk \rightarrow$  & local stack\\
    & $\vdots$\\
@@
 \end{tabular}
-\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been poped to their respective registers. The restore code has been poped of the stack. }
+\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been poped to their respective registers. The restore code has been poped off the stack. }
 \end{figure}
@@
 \label{subsec:example-loc-cap}
+\lau{This text regards the ML-like programs and can be removed when they are.}
 In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1.
@@
 
-Assembly program not using stack. Assume that $r_l \not\in \{\pcreg,r_0 \}$ is a register.
+Assembly program not using stack. Assume that $\mathtt{r_l} \not\in \{\pcreg,r_0 \}$ is a register.
 \begin{verbatim}
@@
   &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \text{ for all } n
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
 \end{align*}
@@
 \begin{lemma}
+  \label{lem:correctness-f2}
   For all $n \in \nats$
@@
     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
-    m & \defeq \hs_{f1} \uplus 
+    \reg & \in \Regs \\
+    m & \defeq \hs_{f2} \uplus 
         \hs_\flag \uplus                
@@
         \ms_{\malloc} \uplus 
-        \ms_{\var{stk}}
+        \ms_{\var{stk}} \uplus
         \ms_{\var{frame}} 
@@
     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \text{ for all } n
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
   \end{align*}
@@
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-  \item $\ms_\stk(a) = 0$ for all $a \in [\start_\stk, \addrend_\stk]$
   \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
@@
   \[
-    (\reg_0\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
   \]
@@
 \end{lemma}
+\begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
+  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
+  \begin{align*}
+    &(\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
+    &\quad(\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
+  \end{align*}
+Where 
+\[
+  m''  \defeq \hs_{f2} \uplus 
+        \hs_\flag \uplus                
+        \ms_\link \uplus 
+        \hs_\adv \uplus 
+        \ms_\malloc \uplus 
+        \ms_\stk' \uplus
+        \ms_{\var{frame}}
+\]
+Only the stack has changed
+\begin{figure}
+  \label{fig:f2-stack-before-adv}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Stack} \\
+\cline{2-2}
+   &         \\
+   & $\vdots$ \\
+\cline{2-2}
+$\start_\stk'':$   & $0$\\
+\cline{2-2}
+$c_\stk',c_\stk'' \rightarrow$   & $c_\stk'$ \\
+\cline{2-2}
+   & $c_{f2}'$ \\
+\cline{2-2}
+$c_{\var{prp}} \rightarrow$   & Restore code \\
+   & $\vdots$ \\
+\cline{2-2}
+   & $1$\\
+\cline{2-2}
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r |c |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $\updatePcPerm{c_\adv}$\\
+\cline{2-2}
+$r_0$  &  $c_{\var{prp}}$ \\
+\cline{2-2}
+$r_1$ & $c_\adv$ \\
+\cline{2-2}
+$r_{\var{priv},1}$ & $w_{p,1}$\\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+& $0$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{Stack and register-file just after jump to $\adv$.}
+\end{figure}
+Where
+\begin{itemize}
+\item The range of authority for 
+  \begin{itemize}
+  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
+  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
+  \end{itemize}
+  the two stack capabilities have the same permissions as $c_\stk$
+\item $c_{f2}'$ points to the return address in f2's code.
+\item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
+\end{itemize}
+Define 
+\begin{enumerate}
+\item 
+  \begin{align*}
+W = \;& [0 \mapsto \iota_{\malloc,0}] \\
+    & [1 \mapsto \iota^\sta(\perma,\ms_{f2})]\\ 
+    & [2 \mapsto \iota^\sta (\perma,\ms_\flag)]\\
+    & [3 \mapsto \iota^\sta (\perma,\ms_\link)]\\
+    & [4 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv} (\perma)]\\
+    & [5 \mapsto \iota^\sta (\temp,\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})] \\
+    & [6 \mapsto \iota^\pwl_{\start_\stk'' + 1, \addrend_\stk} (\temp)]
+  \end{align*}
+\lau{The interesting regions here are region 5 and 6.
+Region 5 governs our private part of the stack whereas region 6 governs the part of the stack we are about to release.
+I don't think it is strictly needed to make them $\temp$ for this example, but in general it is what we wish to do, so we can revoke them upon return.}
+\end{enumerate}
+
+\end{proof}
 
@@
   &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
+  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,1,=,H^{\var{pwl}}) \\ \\
   &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
@@
     \begin{cases}
-      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
+      (v,1,=,H^{\var{nwl}}) & v \neq \revoked \\
       \bot & \text{otherwise}
@@
 
+\subsection{Useful regions}
+Static region used for parts of memory that should not change.
+\begin{align*}
+  \iota^\sta (v,\ms) &= (v,1,=,H^\sta\;\ms)\\
+  H^\sta \; \ms \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
+\end{align*}
+
+\begin{align*}
+  \iota^\cnst (v,n) &= (v,1,=,H^\cnst\;n)\\
+  H^\sta \; n \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \Rightarrow \forall a \in \dom(\ms)\ldotp \ms(a) = n \}
+\end{align*}
+
 \subsection{Lemmas}

### 2016-11-28 local-capability-machine.tex
added comment

@@
 \end{align*}
-
+\lau{With respect to our talk about whether the upper bound should be included in the range of authority or not, I have found one example where it would work better when the upper-bound is not included.
+If we have a stack and call and want to pass the empty part on, then if the unused part of the stack is 0 cells, then we cannot pass anything that ``looks like a stack''.}
 A number of functions and predicates used in the definition of $\sem{-}$ (defined later). Notice all of them are total.

### 2016-11-28 local-capability-machine.tex
Changes based on Lars' mail.

@@
 
-Define the following macros: $\mathtt{restrict}$, $\mathtt{subseg}$, and $\mathtt{lea}$ that does not overwrite the source register. A $\mathtt{store}$ that allows integers to be stored directly.
+Define the following macros: $\mathtt{restrict}$, $\mathtt{subseg}$, and $\mathtt{lea}$ that does not overwrite the source register.
+A $\mathtt{store}$ that allows integers to be stored directly.
+$\mathtt{store}$ requries a register $r_t$ for storage of temporary values to be available.
 \begin{align*}
@@
                    \begin{aligned}[t]
-                     & \move{r_i}{n} \qquad \text{ where $r_i$ unused register}\\
-                     & \store{r}{r_i}
+                     & \move{r_t}{n} \\
+                     & \store{r}{r_t}
                    \end{aligned}
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+2) \\
-    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-2,\mathtt{1f},\mathtt{f1}) \\
-    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+2) \\
+    \var{offset-link-flag} & = 2 \\
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\var{offset-link-flag},\mathtt{1f},\mathtt{f1}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\var{offset-link-flag}) \\
     m & \defeq \hs_{f1} \uplus 
@@
 \begin{align*}
-  &\dom(\hs_{f1}) = [\mathtt{f1}-2,\mathtt{1f}] \\
+  &\dom(\hs_{f1}) = [\mathtt{f1}-\var{offset-link-flag},\mathtt{1f}] \\
   &\dom(\hs_\flag) = [\flag,\flag] \\
@@
 \item $\ms_{\var{ar}}'$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$ (the return address in f1), and activation code.
-\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+2)$ where $b+2$ is the first address of the activation code.
+\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+\var{offset-link-flag})$ where $b+\var{offset-link-flag}$ is the first address of the activation code.
 \end{itemize}
@@
   \begin{itemize}
-  \item Show: $\mapsto \iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.
+  \item Show: $\iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.
   \end{itemize}
@@
   \begin{enumproof}
-    \item Split the memory into the disjoint unions of \ref{f1:m3-disj-union} and show:
+    \item Split the memory into the disjoint unions of \ref{f1:ms-disj-union} and show:
       \begin{enumproof}
@@
           Let $n^{(3)} \leq n''$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n^{(3)}]{\reg}$ be given\\
-          Show: $\npair[n^{(3)}]{(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms)}$\\
-          Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k''$, $m'$ and $\ms_{\var{frame}}$.
-Due to $\heapSat[\ms']{n''}{W'}$ $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
+          Show: $\npair[n^{(3)}]{(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms')} \in \observations(W')$\\
+          Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms' \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k''$, $m'$ and $\ms_{\var{frame}}$.
+Due to $\heapSat[\ms']{n''}{W'}$, $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
 \\
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+2) \\
-    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-2,\mathtt{2f},\mathtt{f2}) \\
+    \var{offset-link-flag} & \defeq 2
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
+    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\var{offset-link-flag},\mathtt{2f},\mathtt{f2}) \\
     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
@@
   \begin{align*}
-    &\dom(\hs_{f2}) = [\mathtt{f2}-2,\mathtt{2f}] \\
+    &\dom(\hs_{f2}) = [\mathtt{f2}-\var{offset-link-flag},\mathtt{2f}] \\
     &\dom(\hs_\flag) = [\flag,\flag] \\

### 2016-11-28 local-capability-machine.tex
Changes to f2 proof

@@
 \]
-Only the stack has changed
-\begin{figure}
-  \label{fig:f2-stack-before-adv}
-  \centering
-  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
-\multicolumn{2}{l}{Stack} \\
-\cline{2-2}
-   &         \\
-   & $\vdots$ \\
-\cline{2-2}
-$\start_\stk'':$   & $0$\\
-\cline{2-2}
-$c_\stk',c_\stk'' \rightarrow$   & $c_\stk'$ \\
-\cline{2-2}
-   & $c_{f2}'$ \\
-\cline{2-2}
-$c_{\var{prp}} \rightarrow$   & Restore code \\
-   & $\vdots$ \\
-\cline{2-2}
-   & $1$\\
-\cline{2-2}
-\end{tabular}
-\hspace{1cm}
-\begin{tabular}{r |c |}
-\multicolumn{2}{r}{Register file} \\
-\cline{2-2}
-$\pcreg$ & $\updatePcPerm{c_\adv}$\\
-\cline{2-2}
-$r_0$  &  $c_{\var{prp}}$ \\
-\cline{2-2}
-$r_1$ & $c_\adv$ \\
-\cline{2-2}
-$r_{\var{priv},1}$ & $w_{p,1}$\\
-\cline{2-2}
-$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
-\cline{2-2}
-& $0$ \\
-\cline{2-2}
-& $\vdots$ \\
-\cline{2-2}
-\end{tabular}
-\caption{Stack and register-file just after jump to $\adv$.}
-\end{figure}
-Where
+Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
 \begin{itemize}
@@
 \end{itemize}
+
+\begin{figure}
+  \label{fig:f2-stack-before-adv}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Stack} \\
+\cline{2-2}
+   &         \\
+   & $\vdots$ \\
+\cline{2-2}
+$\start_\stk'':$   & $0$\\
+\cline{2-2}
+$c_\stk',c_\stk'' \rightarrow$   & $c_\stk'$ \\
+\cline{2-2}
+   & $c_{f2}'$ \\
+\cline{2-2}
+$c_{\var{prp}} \rightarrow$   & Restore code \\
+   & $\vdots$ \\
+\cline{2-2}
+   & $1$\\
+\cline{2-2}
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r |c |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $\updatePcPerm{c_\adv}$\\
+\cline{2-2}
+$r_0$  &  $c_{\var{prp}}$ \\
+\cline{2-2}
+$r_1$ & $c_\adv$ \\
+\cline{2-2}
+$r_{\var{priv},1}$ & $w_{p,1}$\\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+& $0$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{Stack and register-file just after jump to $\adv$.}
+\end{figure}
 Define 

### 2016-11-28 local-capability-machine.tex
Changed semantics of subseg to allow 'invalid' intervals (allowing us to restrict capabilities to not have authority over any addresses).
Moved offset definition to flag/link section

@@
                                                 \quad\text{and in either case $n_i \in \ints$}\\
-                                                \quad\text{and $\start \leq n_1 \leq n_2 \leq \addrend$}\\
+                                                \quad\text{and $\start \leq n_1$ and $n_2 \leq \addrend$}\\
                                                 \quad\text{and $c = (\permp,n_1,n_2,\addr)$}
@@
 The flag table should never be shared between distrusting parties.
+
+We will often want to make room in memory for a linking-table capability and a flag-table capability.
+We therefore define a constant that represents the offset of the actual code of a function caused by these two capabilities:
+\[
+    \var{offset-link-flag} \defeq 2
+\]
 
@@
   \begin{align*}
-    \var{offset-link-flag} & = 2 \\
     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
@@
   \begin{align*}
-    \var{offset-link-flag} & \defeq 2
     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\

### 2016-11-28 local-capability-machine.tex
changed iota^pwl to also allow permanent regions (to avoid issue in readCondition)

@@
   &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,1,=,H^{\var{pwl}}) \\ \\
-  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+  &\iota_{\start,\addrend}^\pwl \defeq \lambda v \ldotp
+    \begin{cases}
+      (v,1,=,H^\pwl (\start,\addrend)) & v \neq \revoked \\
+      \bot & \text{otherwise}
+    \end{cases}
+  &H^{\var{pwl}} : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{pwl}} (\start,\addrend)\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
     \begin{cases}
-      (v,1,=,H^{\var{nwl}}) & v \neq \revoked \\
+      (v,1,=,H^{\var{nwl}} (start,\addrend)) & v \neq \revoked \\
       \bot & \text{otherwise}
@@
 \\ \\
-  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{nwl}} : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
   &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|

### 2016-11-29 local-capability-machine.tex
f2 proof in progress

--- 2016-11-28 local-capability-machine.tex
@@
     &\exists \iota_{\malloc,0} \ldotp \\
-    &\quad (\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))) \; \land \\
+    &\quad (\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s) ([i \mapsto W(i)]) ) \; \land \\
     &\quad \iota_{\malloc,0}.v = \perma \; \land \\
@@
       \begin{enumproof}
-      \item case: $\npair[n']{\ms_\adv} \in \iota_\malloc'.H (\iota_\malloc'.s) (W)$
+      \item case: $\npair[n']{\ms_\malloc} \in \iota_\malloc'.H (\iota_\malloc'.s) (W)$ 
         \begin{enumproof}
@@
         \end{enumproof}
-      \item case: $\npair[n']{\ms_\adv} \in H^\nwl_{\start_\adv,\addrend_\adv} (\start_\adv,\addrend_\adv) W$
+      \item case: $\npair[n']{\ms_\adv} \in H^\nwl_{\start_\adv,\addrend_\adv} 1 W$ \label{lem:f1-adv-mem-sat}
         \begin{enumproof}
@@
   Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
-  \begin{align*}
-    &(\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
-    &\quad(\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
-  \end{align*}
+  \begin{multline*}
+    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
+    (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
+  \end{multline*}
 Where 
@@
 \item $c_{f2}'$ points to the return address in f2's code.
+\item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
 \item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
@@
 \begin{figure}
-  \label{fig:f2-stack-before-adv}
   \centering
@@
 \caption{Stack and register-file just after jump to $\adv$.}
+\label{fig:f2-stack-before-adv}
 \end{figure}
 Define 
-\begin{enumerate}
-\item 
   \begin{align*}
@@
     & [5 \mapsto \iota^\sta (\temp,\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})] \\
-    & [6 \mapsto \iota^\pwl_{\start_\stk'' + 1, \addrend_\stk} (\temp)]
+    & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk} (\temp)]
   \end{align*}
@@
 I don't think it is strictly needed to make them $\temp$ for this example, but in general it is what we wish to do, so we can revoke them upon return.}
-\end{enumerate}
+Show
+\[
+  \npair{(\start_\adv,\addrend_\adv)}\in\readCond{}(\glob)(\iota^\pwl_{\start_\adv,\addrend_\adv})
+\]
+Which amounts to
+\[
+\iota^\nwl_{\start_\adv,\addrend_\adv} (\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv} (\perma)
+\]
+which is indeed the case. Use this to apply the FTLR to get 
+\begin{enumproof}
+\item $\npair{c_\adv} \in \stder(W)$
+\end{enumproof}
+Define
+\[
+  \ms  \defeq \hs_{f2} \uplus 
+        \hs_\flag \uplus                
+        \ms_\link \uplus 
+        \hs_\adv \uplus 
+        \ms_\malloc \uplus 
+        \ms_\stk'
+\]
+($m''$ without the frame) and show
+\begin{enumproof}[resume]
+\item $\heapSat[\ms]{n}{W}$ \lau{The argument here is a bit informal}
+  \begin{itemize}
+  \item For $\hs_{f2}$, $\hs_\flag$, and $\ms_\link$ the regions $\iota^\sta(\perma,\ms_{f2})$,$\iota^\sta (\perma,\ms_\flag)$, and $\iota^\sta (\perma,\ms_\link)$ are designed to respectively accept each of these memory segments.
+  \item $\hs_\adv$, showing this case is similar to item \ref{lem:f1-adv-mem-sat} in the proof of Lemma~\ref{lem:correctness-f1}.
+  \item $\ms_\malloc$, this case follows from the specification of malloc and the assumption that $\iota_{\malloc,0}$ models $\ms_\malloc$.    
+  \item $\ms_\stk'$, here we need to split the memory segment in two and for $n' < n$ argue
+    \begin{itemize}
+    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$:\\ This region matches only this memory segment.
+    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk'', \addrend_\stk]}} \in H_{\start_\stk'', \addrend_\stk}(1)(W)$:\\ Show $\forall a \in [\start_\stk'', \addrend_\stk] \ldotp \ms_\stk'|_{[\start_\stk'', \addrend_\stk]}(a) \in \stdvr(W)$, trivially true since these memory cells contain 0.
+    \end{itemize}
+  \end{itemize}
+\item $\npair{\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}} \in \stdrr(W)$
+  \begin{enumproof}
+    \item case $\npair{c_\stk''} \in \stdvr(W)$: Show:
+      \begin{enumproof}
+        \item $\npair{(\start_\stk'', \addrend_\stk)} \in \readCond{}(\local)(W)$ \\
+          This amounts to showing 
+\[
+  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+\]
+which is indeed the case.
+        \item $\npair{(\start_\stk'', \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$\\
+This amounts to showing 
+\[
+  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+\]
+which is indeed the case.
+        \item $\npair{(\rwlx,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$ \label{f2:rwlx-exec-case} \\
+We will show this case in some detail, the others are similar. \\
+Let $n' < n$, $W' \futurewk W$, and $a \in [\start_\stk'', \addrend_\stk]$ be given. Show
+\[
+  \npair[n']{((\rwlx,\local),\start_\stk'', \addrend_\stk,a)} \in \stder(W')
+\]
+If we can apply the FTLR at this point, then we are done. SFTS:
+\begin{enumproof}
+  \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \readCond{}(\local)(W')$\\
+this amounts to showing
+\[
+  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+\]
+which is indeed the case.
+  \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \writeCond{}(\local)(W')$ \\
+this amounts to showing
+\[
+  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+\]
+which is indeed the case.
+\end{enumproof}
+        \item $\npair{(\rwx,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+Similar to case~\ref{f2:rwlx-exec-case}
+        \item $\npair{(\exec,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+Similar to case~\ref{f2:rwlx-exec-case}
+      \end{enumproof}
+    \item case $\npair{c_\adv} \in \stdvr(W)$:\\
+     Follows from the FTLR.
+    \item case $\npair{c_{\var{prp}}} \in \stdvr(W)$:\\
+      Let $n' < n$ and $W' \future W$ be given and show
+      \[
+        \npair[n']{\updatePcPerm{c_{\var{prp}}}} \in \stder(W')
+      \]
+      to this en let $n'' \leq n'$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n'']{\reg'} \in \stdrr$ be given and show
+      \[
+        \npair[n'']{(\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms')} \in \observations(W')
+      \]
+      Let $i \leq n''$, $m''$ and $\ms_f$ be given and assume
+      \[
+        (\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms'\uplus\ms_f) \step[i]
+        (\halted,m'')
+      \]
+      (From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack).
+The restoration code remains unaltered, and so does the data on the stack, so the execution continues by poping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$.
+Continuing in $f2$, the restore code is poped off the stack which is the last of the $\mathtt{scall}$ instruction.
+Now the top word of the stack is poped to $r_1$, and thanks to the world satisfaction, we know that this word is 1.
+The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.)
+
+Use $\revokeTemp(W') \futurestr W'$. From the private-future memory satisfaction lemma (Lemma~\ref{lem:priv-mono-like}) get $\ms''$ and $\ms_f$ such that
+
+    \item case $r\not\in \{\pcreg,r_0,r_1,r_\stk\}$: $\reg_0(r) = 0$. Trivial.
+  \end{enumproof}
+\end{enumproof}
+to get
 
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,1,=,H^{\var{pwl}}) \\ \\
-  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+  &\iota_{\start,\addrend}^\pwl : \Views \fun \Regions\\
+  &\iota_{\start,\addrend}^\pwl \defeq \lambda v \ldotp
+    \begin{cases}
+      (v,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
+      \bot & \text{otherwise}
+    \end{cases} \\
+  &H^\pwl : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
+  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
 \end{align*}
+\lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
 Revoking all temporary regions:
@@
     \begin{cases}
-      (v,1,=,H^{\var{nwl}}) & v \neq \revoked \\
+      (v,1,=,H^\nwl_{start,\addrend}) & v \neq \revoked \\
       \bot & \text{otherwise}
@@
 \\ \\
-  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
-  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^\nwl : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
                         & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-                        & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
+                        & \;\qquad W(r)\nsubeq[n-1] \iota^\pwl_{\start',\addrend'}(W(r).v) \}
             \end{aligned}
@@
              & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \}
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
              \union \\
@@
              & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \land\\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
              & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
@@
 \begin{lemma}[Private monotonicity-like property]
+\label{lem:priv-mono-like}
   \begin{align*}
@@
     & \qquad \Rightarrow \npair{\stdcap[(\perm,\glob)]} \in \stdvr(\wrev{W}) 
+  \end{align*}
+\end{lemma}
+
+\subsubsection{Standard regions}
+\begin{lemma}
+  \begin{align*}
+    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
+    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubeq \iota^\pwl_{\start,\addrend}(v)
   \end{align*}

### 2016-11-29 local-capability-machine.tex
Added 'nsubsetsim' (almost like n-subset) and changed the standard islands.

@@
 \newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
+
+\newcommand\subsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\subset$}\cr
+  \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
+\newcommand\supsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\supset$}\cr
+  \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
+\newcommand{\nsubsim}[1][n]{\stackrel{\tiny{#1}}{\subsetsim}}
+\newcommand{\nsupsim}[1][n]{\stackrel{\tiny{#1}}{\supsetsim}}
+
 \newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
 \newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
+
 \newcommand{\union}{\mathbin{\cup}}
@@
 \section{Domains and Notation}
+
 \begin{align*}
@@
 \item $W = [0 \mapsto \iota_\malloc']
-           [1 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv}(\perma)]
+           [1 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
            [2 \mapsto \iota_{f1}]
@@
 \end{itemize}
+\lau{Need to change this to use the new standard regions}
 where
@@
   \begin{itemize}
-  \item Show: $\iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.
+  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}$.
   \end{itemize}
@@
           \item $a = \start_\adv$: show $((\readonly,\glob),\link,\link+1,\link) \in \stdvr(W)$\\
-           SFTS $\iota_{\link} \nsubeq[n'-1] \iota^\pwl_{\link,\link+1} (\perma)$\\
+           SFTS $\iota_{\link} \nsubsim[n'-1] \iota^\pwl_{\link,\link+1}$\\
            $\iota_{\link}$ accepts only $\ms_\link$, so show:
@@
     & [3 \mapsto \iota^\sta (\perma,\ms_\link)]\\
-    & [4 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv} (\perma)]\\
+    & [4 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
     & [5 \mapsto \iota^\sta (\temp,\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})] \\
-    & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk} (\temp)]
+    & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk}]
   \end{align*}
@@
 \[
-\iota^\nwl_{\start_\adv,\addrend_\adv} (\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv} (\perma)
+\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}
 \]
@@
 \[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
 \]
@@
 \[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsupsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
 \]
@@
 \[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
 \]
@@
 \[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
+  \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsupsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
 \]
@@
   &\iota_{\start,\addrend}^\pwl : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq \lambda v \ldotp
-    \begin{cases}
-      (v,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
-      \bot & \text{otherwise}
-    \end{cases} \\
+  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
+\\
   &H^\pwl : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^{\var{nwl}} : \Views \parfun \Regions \\
-  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp
-    \begin{cases}
-      (v,1,=,H^\nwl_{start,\addrend}) & v \neq \revoked \\
-      \bot & \text{otherwise}
-\end{cases}
-\\ \\
+  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
+  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,H^\nwl_{start,\addrend}) \\
+\\
+  &\iota_{\start,\addrend}^{\nwl,p} : \Regions \\
+  &\iota_{\start,\addrend}^{\nwl,p} \defeq (\perma,1,=,H^\nwl_{start,\addrend})\\
+\\
   &H^\nwl : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
@@
 
+We need a notion of subset between regions that is almost $n$-subset, but not quite.
+The only difference is that the view part of a region is diregarded.
+Define ``semi $n$-subset'' and ``semi $n$-supset'' as:
+\lau{ Using normal the canonical $n$-subset gave some technical problems in the proofs.
+For instance, it was not possible to have a global read capability for a region governed by a permanent standard region.
+}
+\begin{mathpar}
+  \inferrule{(s,H) = (s',H) \\
+              \forall \hat{W} \ldotp H \; s \; \hat{W} \nsubeq H' \; s' \; \hat{W} }
+            { (v,s,\phi,H) \nsubsim (v',s',\phi',H')}
+\end{mathpar}
+
 \begin{align*}
@@
               & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-              & \;\qquad W(r)\nsupeq[n-1] \iota_{\start',\addrend'}(W(r).v) \}
-            \end{aligned}
-\end{align*}
-
-\begin{align*}
-    & \readCond{}(\gl)(W) =  \\
-    & \quad \begin{aligned}[t]
-              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-                        & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-                        & \;\qquad W(r)\nsubeq[n-1] \iota^\pwl_{\start',\addrend'}(W(r).v) \}
+              & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \}
             \end{aligned}
@@
              & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \}
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
              \union \\
@@
              & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \land\\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
              & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
@@
     & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubeq \iota^\pwl_{\start,\addrend}(v)
+    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
   \end{align*}

### 2016-11-29 local-capability-machine.tex
further changes related to the new standard regions
f2 proof

@@
 \end{tabular}
-\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been poped to their respective registers. The restore code has been poped off the stack. }
+\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been popped to their respective registers. The restore code has been popped off the stack. }
 \end{figure}
@@
 \begin{enumproof}
-\item $\npair{c_\adv} \in \stder(W)$
+\item $\npair{c_\adv} \in \stder(W)$ \label{lem:f2-adv-er}
 \end{enumproof}
@@
 \begin{enumproof}[resume]
-\item $\heapSat[\ms]{n}{W}$ \lau{The argument here is a bit informal}
+\item $\heapSat[\ms]{n}{W}$ \label{lem:f2-mem-sat}\lau{The argument here is a bit informal}
   \begin{itemize}
@@
   \end{itemize}
-\item $\npair{\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}} \in \stdrr(W)$
+\item $\npair{\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}} \in \stdrr(W)$ \label{lem:f2-reg-rel}
   \begin{enumproof}
@@
       \]
-      (From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack).
-The restoration code remains unaltered, and so does the data on the stack, so the execution continues by poping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$.
-Continuing in $f2$, the restore code is poped off the stack which is the last of the $\mathtt{scall}$ instruction.
-Now the top word of the stack is poped to $r_1$, and thanks to the world satisfaction, we know that this word is 1.
-The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.)
-
-Use $\revokeTemp(W') \futurestr W'$. From the private-future memory satisfaction lemma (Lemma~\ref{lem:priv-mono-like}) get $\ms''$ and $\ms_f$ such that
-
+
+      From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack).
+The restoration code remains unaltered, and so does the data on the stack, so the execution continues by popping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$.
+Continuing in $f2$, the restore code is popped off the stack which is the last of the $\mathtt{scall}$ instruction.
+Now the top word of the stack is popped to $r_1$, and thanks to the world satisfaction, we know that this word is 1.
+The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.
+
+All in all, there are no changes to the memory, so we can conclude
+\[
+  m'' = \ms' \uplus \ms_f
+\]
+From Lemma~\ref{lem:priv-mono-like}, we can get $\ms''$ and $\ms_r$ such that
+\begin{itemize}
+\item $\ms' = \ms'' \uplus \ms_r$
+\item $\heapSat[\ms'']{n}{\revokeTemp{W}}$
+\end{itemize}
+Now downwards closure of memory segment satisfaction gives us the desired result.
     \item case $r\not\in \{\pcreg,r_0,r_1,r_\stk\}$: $\reg_0(r) = 0$. Trivial.
@@
 \end{enumproof}
-to get
-
+From \ref{lem:f2-adv-er}, \ref{lem:f2-mem-sat}, and \ref{lem:f2-reg-rel} it follows that
+\[
+  \npair{(\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}\update{\pcreg}{c_\adv},\ms)} \in \observations(W)
+\]
+Now using $i$, the number of steps we need to halt, $\ms_f$ as the frame, $m'$ as the resulting memory and the assumption that the execution terminates successfuly gives $W' \futurestr W$, $\ms'$, and $\ms_r$ such that
+\begin{itemize}
+\item $m' = \ms' \uplus\ms_r \uplus \ms_f$
+\item $\heapSat[\ms']{n-i}{W'}$
+\end{itemize}
+The permanent region $\iota^\sta (\perma,\ms_\flag)$ ensures that $\ms_\flag$ is unchanged, so
+\[
+  m'(\flag) = 0
+\]
 \end{proof}
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^\pwl : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
+  &\iota_{\start,\addrend}^\pwl : \Regions\\
+  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,H^\pwl_{\start,\addrend}) \\
 \\

### 2016-11-30 local-capability-machine.tex
added f3 program (next to be proven)
Small remark.

--- 2016-11-29 local-capability-machine.tex
@@
 \newcommand{\pwl}{\var{pwl}}
-\newcommand{\sta}{\var{sta}}
-\newcommand{\cnst}{\var{cnst}}
 
@@
 \newcommand{\Heaps}{\plaindom{Mem}}
-\newcommand{\Mems}{\Heaps}
 \newcommand{\HeapSegments}{\plaindom{MemSegment}}
-\newcommand{\MemSegments}{\HeapSegments}
 \newcommand{\Confs}{\plaindom{Conf}}
@@
 \ExecConfs  &::= \Regs \times \Heaps \\
-\Globals & ::= \{\glob, \local \} \\
 \Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
-\Confs &::= \ExecConfs + \{\failed \} + \halted \times \ExecConfs \\
-\HeapSegments &::= \Addrs \parfun \Words
+\Confs &::= \ExecConfs + \{\failed, \halted \times \ExecConfs\} \\
+\HeapSegments &::= \Addrs \parfun \Words\\
+\Globals & ::= \{\glob, \local \}
 \end{align*}
 Local capabilities have been added by adding a new domain $\Globals$ which represents whether a capability is local or global.
-There are two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities.
+Two new permissions $\readwritel$ and $\rwlx$ that permits writing local capabilities.
 They are otherwise the same as their non-''permit write local'' counterparts.
 
+The $\glob$ and $\local$ are ordered as shown in Figure~\ref{fig:glob-hier}.
 \begin{figure}[!h]
@@
 \hline
-  $\readwrite$ & Read-write. Grants read and write permission. Storage of local capabilities prohibited. \\
+  $\readwrite$ & Read-write. Grants read and write permissions. Storage of local capabilities prohibited. \\
 \hline
-  $\readwritel$ & Read-write, permit write local. Grants read and write permission. Storage of local capabilities possible. \\
+  $\readwritel$ & read-write, permit write local. Grants read and write permissions. Storage of local capabilities granted. \\
 \hline
@@
 \hline
-  $\entry$ & Enter permission. This permission grants no access, but when jumped to, it will turn into an $\exec$ permission.\\
+  $\entry$ & Entry permission. This permission grants no access, but when jumped to, it Will turn into an $\exec$ permission.\\
 \hline
-  $\rwx$ & Read-write-execute permission. Grants read, write, and execute permissions. Storage of local capabilities prohibited. \\
+  $\rwx$ & Read, write, execute permission. Grants read, write, and execute permissions. Storage of local capabilities prohibited. \\
 \hline
-  $\rwlx$ & Read-write-execute, permit write local. Grants read, write, and execute permissions. Storage of local capabilities possible.
+  $\rwlx$ & Read, write, execute permission with permit write local. Grants read, write, and execute permissions. Storage of local capabilities possible.
 \end{tabular}
@@
 \Phi    &\in& \ExecConfs \\
-m, \memheap&\in& \Heaps \\
+\memheap&\in& \Heaps \\
 \memreg &\in& \Regs \\
@@
 n       &\in& \ints\\
-\ms     &\in& \MemSegments
 \end{array}
 \]
-Words and instructions:
+Further definitions:
 \[
@@
 \subsection*{Semantics}
-Assume a $\decode$ function that decodes words to instructions:
+Assume a $\decode$ function that decodes integer to instructions:
 \begin{align*}
@@
 \end{align*}
-\lau{With respect to our talk about whether the upper bound should be included in the range of authority or not, I have found one example where it would work better when the upper-bound is not included.
-If we have a stack and call and want to pass the empty part on, then if the unused part of the stack is 0 cells, then we cannot pass anything that ``looks like a stack''.}
+
 A number of functions and predicates used in the definition of $\sem{-}$ (defined later). Notice all of them are total.
@@
                                                    \quad\text{and in either case $n \in \ints$}\\
-                                                   \quad\text{and $\decodePermPair{n}\sqsubseteq \permp$}\\
-                                                   \quad\text{and $c = (\decodePermPair{n},\start,\addrend,\addr)$}
+                                                   \quad\text{and $\var{newPermPair} = \decodePermPair(n)$}\\
+                                                   \quad\text{and $\var{newPermPair} \sqsubseteq \permp$}\\
+                                                   \quad\text{and $c = (\var{newPermPair},\start,\addrend,\addr)$}
                                                  \end{array}\\
@@
 \end{align*}
-\lau{Do we need other arithmetic operators than plus?}
 \begin{align*}
@@
                                               \begin{array}[t]{l}
-                                                \text{if $\memreg(r) = \stdcap[\permp]$} \\
+                                                \text{if $\memreg{r} = \stdcap[\permp]$} \\
                                                 \quad\text{and for $i \in \{1,2\}$}\\
@@
                                                 \quad\text{and in either case $n_i \in \ints$}\\
-                                                \quad\text{and $\start \leq n_1$ and $n_2 \leq \addrend$}\\
-                                                \quad\text{and $c = (\permp,n_1,n_2,\addr)$}
+                                                \quad\text{and $\start \leq n_1 \leq n_2 \leq \addrend$}\\
+                                                \quad\text{and $c = (\permp,n_2,n_2,\addr)$}
                                               \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$}
+                              \text{if $\memreg{r_2} = \stdcap$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg(r_2) = ((\_,\_),\start,\_,\_)$}
+                              \text{if $\memreg{r_2} = \stdcap$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg(r_2) = ((\_,\_),\_,\addrend,\_)$}
+                              \text{if $\memreg{r_2} = \stdcap$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg(r_2) = ((\perm,\_),\_,\_,\_)$}
+                              \text{if $\memreg{r_2} = \stdcap$}
                             \end{array} \\
@@
                             \begin{array}[t]{l}
-                              \text{if $\memreg(r_2) = ((\_,\gl),\_,\_,\_)$}
+                              \text{if $\memreg{r_2} = \stdcap$}
                             \end{array} \\
@@
                           \begin{cases}
-                            \stdUpdatePc{\updateReg{r_1}{1}} & \text{if $\memreg(\rv) \in \Caps$ } \\
+                            \stdUpdatePc{\updateReg{r_1}{1}} & \text{if $\memreg{\rv} = \stdcap$} \\
                             \stdUpdatePc{\updateReg{r_1}{0}} & \text{otherwise} 
@@
 
-After adding local capabilities, most of the definitions have changed slightly to take into account the new form the capabilities take.
-The only two instructions that have undergone enough change to deserve to be mentioned are $\mathtt{store}$ and $\mathtt{restrict}$.
-
-$\mathtt{store}$ has the same conditions as before, but if what we try to write is a local capability, then it needs to be a write access that permits write local.
-
-$\mathtt{restrict}$ looks like before, but now it uses a ``permission pair'' which is the access permission and the locality.
-Whether a $\mathtt{restrict}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above (it is the point-wise ordering of the pair).
-
-Define the following macros: $\mathtt{restrict}$, $\mathtt{subseg}$, and $\mathtt{lea}$ that does not overwrite the source register.
-A $\mathtt{store}$ that allows integers to be stored directly.
-$\mathtt{store}$ requries a register $r_t$ for storage of temporary values to be available.
+After adding local capabilities, most of the definitions have changed slightly to take into the account new form the capabilities take.
+The only two instructions that have undergone enough change to deserve to be mentioned are $\store{}$ and $\restrict{}{}{}$.
+
+$\store{}$ has the same conditions as before, but if what it tries to write is a capability, then it needs to be a write access that permits write local.
+
+$\restrict{}{}{}$ looks like before, but now it uses a ``permission pair'' which is the access permission and whether it is local or not.
+Whether a $\restrict{}{}{}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above.
+
+Define the following simple macros:
+
 \begin{align*}
@@
                    \begin{aligned}[t]
-                     & \move{r_t}{n} \\
-                     & \store{r}{r_t}
+                     & \move{r_i}{n} \qquad \text{ where $r_i$ unused register}\\
+                     & \store{r}{r_i}
                    \end{aligned}
@@
   \begin{aligned}
-    &\exists \iota_{\malloc,0} \ldotp \\
-    &\quad (\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s) ([i \mapsto W(i)]) ) \; \land \\
-    &\quad \iota_{\malloc,0}.v = \perma \; \land \\
-    &\quad (\forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
+    & \exists \iota_{\malloc,0} \ldotp \\
+    & \quad \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
     &\qquad \forall i, n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \\
@@
     &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-    &\qquad \qquad \quad \exists j \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
     &\qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
@@
     &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-    &\qquad \qquad \qquad \size - 1 = e'-b' )
+    &\qquad \qquad \qquad \size - 1 = e'-b' 
   \end{aligned}
 \]
+where $\iota_{\malloc,0}$ is the region that initially governs malloc 
+\[
+\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s ([i \mapsto W(i)]))
+\]
+and
+\[
+  \iota_{\malloc,0}.v = \perma
+\]
+
 \end{specification}
 In the specification above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
+\lau{should it be a public or private future world in the premise?}
 
@@
 In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
+\dominique{how is the return capability constructed: not on the stack, I presume, so  means mallocing a block of memory for the activation record?}
+\lau{See the new addition below - when we don't have a stack, then we have no need to ``save'' local capabilities which means that we can have the return code as part of the program.
+The only thing we have to do is to have the capability point to the return address and make the capability into an entry capability.
+}
+\lau{On further thought, conveniently the examples did not seem to need an activation record, but if we want \texttt{call} to work more generally speaking, then it needs to setup an activation record.
+I have changed the description of \texttt{call}, so it now uses an activation record.
+}
 \dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
@@
 In order to make capabilities to trusted code (and possibly untrusted code) available, we assume that some sort of linker has made these available.
-This is done in the following way: For every function, the first memory cell the capability for that function governs contains a capability for the linking table.
-Each function name in a program corresponds to an offset in the table, e.g., \texttt{malloc} could be at offset 0.
-When a name is used in a program, it indicates what entry from the linking table to pick.
-The table should always be accessible by taking a copy of the capability in the $\pcreg$-register and adjusting it to point to the first cell it governs.
-
-The capability linking table can be shared between multiple functions that are linked to the same capabiblities as it is accessed through read-only capabilities.
-
-\subsection{Flag table}
+This is done in the following way: For every function, the first memory cell the capability for that function governs contains a capability for the list in question.
+Each function name in a program corresponds to an offset in the table, e.g., \texttt{malloc}.
+When a name is used in a program, it indicates what entry to pick from the list.
+The list should always be accessible by taking a copy of the capability in the $\pcreg$-register and adjusting it to point to the first cell it governs.
+
+The capability link list can be shared between multiple functions as it is accessed through a read-only capability.
+
+\subsection{Flag list}
 A function may use flags to signal failure.
-We use the convention that a flag table is available in the second memory cell of a functions code (so just after the linking table).
-The flag table is accessed through a read-write capability and initially it contains all zero.
-Like the linking table, each entry is associated with a name which may appear in the macros.
-
-The flag table should never be shared between distrusting parties.
-
-We will often want to make room in memory for a linking-table capability and a flag-table capability.
-We therefore define a constant that represents the offset of the actual code of a function caused by these two capabilities:
-\[
-    \var{offset-link-flag} \defeq 2
-\]
-
-\subsection{Macro definitions}
+We use the convention that a flag list is available in the second memory cell of a functions code (so just after the linking table).
+The flag list is accessed through a read-write capability and initially it contains all zero.
+Like the linking list, each entry is associated with a name which may appear in the macros.
+
+The flag list should never be shared between distrusting parties.
+
 \begin{description}
-\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline
-$\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
+\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
   \begin{itemize}
   \item Set up activation record
-  \item Create local enter capability for activation (protected return pointer)
+  \item Create local enter capability for activation
   \item Clear unused registers
@@
 \begin{description}
-  \item [Set up activation record]\forcenewline
+  \item [Set up activation record]
     \begin{itemize}
@@
     \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
-    \item Adjust a copy of the current pc to point to the return address in code and save it to the activation record.
+    \item Adjust a copy of the current pc to point to the return address and save it to the activation record.
     \item Write the activation code to the activation record.
   \end{itemize}
-  \item [Create local enter capability for activation] Adjust the capability for the activation record to point to the beginning of the activation record and restrict it to a local enter-capability.
+  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability.
 Place this capability in $r_0$.
@@
 %    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
-    \item Load the return capability to $\pcreg$
+    \item Load the return address to $\pcreg$
     \end{itemize}
 \end{description}
-\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register $r$.
+\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
 
@@
 If they are equal, then execution continues.
-If they are unequal, then the assertion flag named $\var{flag}$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
+If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
 \item[\texttt{cclear $r$}] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
@@
 \begin{itemize}
-\item \texttt{call} will fail if we have local capabilities in one of the registers of the ``private'' register list as it relies on a capability returned by malloc which will not be permit-write-local.
+\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local.
 Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
@@
 \subsection{Stack}
-Some programs will assume access to a stack which will be in part indicated by the program macros but also in the correctness lemma.
+Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma.
 The stack is accessed through a local $\rwlx$-capability.
@@
 
-The stack resides entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
+The stack will reside entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
 
@@
 When not in the middle of a push or a pop, the stack capability points to the top word of the stack.
-For an empty stack, the stack capability points to the address just below of the range of authority for the stack capability.
-
-The stack grows upwards
+For an empty stack, the stack capability points to the address just outside of the range of authority for the stack capability.
 
 \begin{description}
-\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the address of the stack capability by one and storing the word through the stack capability.
-\item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, and decrementing the address of the stack capability.
-\item[\texttt{scall} \forcenewline
-$f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
+\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the stack capability by one and storing the word through said capability.\footnote{The stack grows upwards.}
+\item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
+\item[\texttt{scall} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
 This call assumes $r_{\var{stk}}$ contains a stack capability.
@@
   \begin{itemize}
-  \item Push the restore code to the stack.
+  \item Push restore code to the stack.
   \item Push ``private'' registers to the stack.
   \item Push return address capability 
-  \item Push stack capability
+  \item Push stack pointer
   \item Create protected return pointer
-  \item Restrict stack capability to unused part 
-  \item Clear the part of the stack we release control over
+  \item Restrict stack capability
+  \item Clear the stack we release control over
   \item Clear unused registers
   \item Jump
-  \item Upon return: Run the on stack restore code
-  \item Return address in caller-code:  Restore ``private'' state
+  \item Upon return: Run restore code
+  \item Return address: 
+  \item Restore ``private'' state
   \end{itemize}
@@
 \begin{description}
-  \item [Push the restore code to the stack]
-    Push the restore code to the stack (described later).
-This code needs to be on the stack to make sure the stack capability can be restored.
-We keep the restore code on the stack minimal.
-The caller code does the rest of the restoration.
+  \item [Push restore code to the stack]
+    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
   \item [Push ``private'' registers to the stack]
@@
     Push a capability for the return address (in the memory) to the stack.
-  \item [Push stack capability]
-    Push the full stack capability to the stack.
+  \item [Push stack pointer]
+    Push the full stack pointer to the stack.
   \item [Create protected return pointer]
@@
 Restrict it to a local enter-capability and put it in $r_0$.
-\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?} \lau{It is also used when we want to call code for the first time, so calling it a protected return pointer is more of a conceptul thing.}
-  \item [Restrict stack capability to unused part]
+\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?}
+  \item [Restrict stack capability]
     Make the stack capability only govern the unused part.
-  \item [Clear the part of the stack we release control over]
+  \item [Clear the stack we release control over]
     Store 0 to all the memory cells the restricted stack pointer has authority over.
@@
   \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
-  \item [Run the on stack restore code]
-    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter (the return address in caller-code) from the stack to $\pcreg$.
-  \item [Return address in caller-code: Restore ``private'' state] \forcenewline
+  \item [Restore code]
+    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
+  \item [Return Address:] \forcenewline
+    \begin{description}
+    \item [Restore ``private'' state] \forcenewline
       \begin{itemize}
@@
       \item Pop the restore code of the stack
-      \end{itemize}\
+      \end{itemize}
+  \end{description}
 \end{description}
@@
 It is also possible to use a combination of \texttt{scall} and \texttt{call}, but when \texttt{call} is used, then we have no way to store the stack, so we cannot use \texttt{scall} after that.
-\item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared before it is passed to an untrusted party.
+\item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared upon return.
 \item As a rule of thumb: If you receive a stack from an untrusted source, then you need to check that it is a local $\rwlx$-capability and clear it!
@@
    &  \\
+   & $\vdots$ \\
+\cline{2-2}
+   & ?\\
+\cline{2-2}
+   & 0 \\
+\cline{2-2}
    & $\vdots$\\
 \cline{2-2}
-   & ? \\
-\cline{2-2}
-$ \rightarrow$  & $c_0$ \\
-\cline{2-2}
-   & $w_{p,1}$ \\
-\cline{2-2}
-   & $\vdots$ \\
-\cline{2-2}
-   & $w_{p,m}$ \\
-\cline{2-2}
-   & Restore code \\
-   & $\vdots$\\
-\cline{2-2}
- $c_\stk \rightarrow$  & local stack\\
+   & 0 \\
+\cline{2-2}
+$c_{\var{stk}} \rightarrow$   & local stack\\
    & $\vdots$\\
@@
 \end{tabular}
-\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been poped to their respective registers. The restore code has been poped off the stack. }
+\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been poped to their respective registers. The restore code has been poped of the stack. }
 \end{figure}
@@
 \label{subsec:example-loc-cap}
-\lau{This text regards the ML-like programs and can be removed when they are.}
 In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1.
@@
 
-Assembly program not using stack. Assume that $\mathtt{r_l} \not\in \{\pcreg,r_0 \}$ is a register.
+Assembly program not using stack. Assume that $r_l \not\in \{\pcreg,r_0 \}$ is a register.
 \begin{verbatim}
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
-    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\var{offset-link-flag},\mathtt{1f},\mathtt{f1}) \\
-    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\var{offset-link-flag}) \\
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+2) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-2,\mathtt{1f},\mathtt{f1}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+2) \\
     m & \defeq \hs_{f1} \uplus 
@@
 \begin{align*}
-  &\dom(\hs_{f1}) = [\mathtt{f1}-\var{offset-link-flag},\mathtt{1f}] \\
+  &\dom(\hs_{f1}) = [\mathtt{f1}-2,\mathtt{1f}] \\
   &\dom(\hs_\flag) = [\flag,\flag] \\
@@
   &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \text{ for all } n
 \end{align*}
@@
 \item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f1}$ contains the code of $f1$.
+\dominique{why not write this in symbols?}
 \item $\ms_\flag = [flag \mapsto 0]$
@@
 \item $\ms_{\var{ar}}'$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$ (the return address in f1), and activation code.
-\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+\var{offset-link-flag})$ where $b+\var{offset-link-flag}$ is the first address of the activation code.
+\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+2)$ where $b+2$ is the first address of the activation code.
 \end{itemize}
@@
   \begin{itemize}
-  \item Show: $\iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.
+  \item Show: $\mapsto \iota^\nwl_{\start_\adv,\addrend_\adv}(\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}(\perma)$.
   \end{itemize}
@@
   \begin{enumproof}
-    \item Split the memory into the disjoint unions of \ref{f1:ms-disj-union} and show:
+    \item Split the memory into the disjoint unions of \ref{f1:m3-disj-union} and show:
       \begin{enumproof}
-      \item case: $\npair[n']{\ms_\malloc} \in \iota_\malloc'.H (\iota_\malloc'.s) (W)$ 
+      \item case: $\npair[n']{\ms_\adv} \in \iota_\malloc'.H (\iota_\malloc'.s) (W)$
         \begin{enumproof}
@@
         \end{enumproof}
-      \item case: $\npair[n']{\ms_\adv} \in H^\nwl_{\start_\adv,\addrend_\adv} 1 W$ \label{lem:f1-adv-mem-sat}
+      \item case: $\npair[n']{\ms_\adv} \in H^\nwl_{\start_\adv,\addrend_\adv} (\start_\adv,\addrend_\adv) W$
         \begin{enumproof}
@@
           Let $n^{(3)} \leq n''$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n^{(3)}]{\reg}$ be given\\
-          Show: $\npair[n^{(3)}]{(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms')} \in \observations(W')$\\
-          Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms' \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k''$, $m'$ and $\ms_{\var{frame}}$.
-Due to $\heapSat[\ms']{n''}{W'}$, $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
+          Show: $\npair[n^{(3)}]{(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms)}$\\
+          Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k''$, $m'$ and $\ms_{\var{frame}}$.
+Due to $\heapSat[\ms']{n''}{W'}$ $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
 \\
@@
 \begin{lemma}
-  \label{lem:correctness-f2}
    For all $n \in \nats$
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
-    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\var{offset-link-flag},\mathtt{2f},\mathtt{f2}) \\
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+2) \\
+    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-2,\mathtt{2f},\mathtt{f2}) \\
     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
-    \reg & \in \Regs \\
-    m & \defeq \hs_{f2} \uplus 
+    m & \defeq \hs_{f1} \uplus 
         \hs_\flag \uplus                
@@
         \ms_{\malloc} \uplus 
-        \ms_{\var{stk}} \uplus
+        \ms_{\var{stk}}
         \ms_{\var{frame}} 
@@
   \begin{align*}
-    &\dom(\hs_{f2}) = [\mathtt{f2}-\var{offset-link-flag},\mathtt{2f}] \\
+    &\dom(\hs_{f2}) = [\mathtt{f2}-2,\mathtt{2f}] \\
     &\dom(\hs_\flag) = [\flag,\flag] \\
@@
     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \text{ for all } n
   \end{align*}
@@
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+  \item $\ms_\stk(a) = 0$ for all $a \in [\start_\stk, \addrend_\stk]$
   \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
@@
   \[
-    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+    (\reg_0\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
   \]
@@
 \end{lemma}
-\begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
-  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
-  \begin{multline*}
-    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
-    (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
-  \end{multline*}
-Where 
-\[
-  m''  \defeq \hs_{f2} \uplus 
-        \hs_\flag \uplus                
-        \ms_\link \uplus 
-        \hs_\adv \uplus 
-        \ms_\malloc \uplus 
-        \ms_\stk' \uplus
-        \ms_{\var{frame}}
-\]
-Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
-\begin{itemize}
-\item The range of authority for 
-  \begin{itemize}
-  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
-  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
-  \end{itemize}
-  the two stack capabilities have the same permissions as $c_\stk$
-\item $c_{f2}'$ points to the return address in f2's code.
-\item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
-\item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
-\end{itemize}
-
-\begin{figure}
-  \centering
-  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
-\multicolumn{2}{l}{Stack} \\
-\cline{2-2}
-   &         \\
-   & $\vdots$ \\
-\cline{2-2}
-$\start_\stk'':$   & $0$\\
-\cline{2-2}
-$c_\stk',c_\stk'' \rightarrow$   & $c_\stk'$ \\
-\cline{2-2}
-   & $c_{f2}'$ \\
-\cline{2-2}
-$c_{\var{prp}} \rightarrow$   & Restore code \\
-   & $\vdots$ \\
-\cline{2-2}
-   & $1$\\
-\cline{2-2}
-\end{tabular}
-\hspace{1cm}
-\begin{tabular}{r |c |}
-\multicolumn{2}{r}{Register file} \\
-\cline{2-2}
-$\pcreg$ & $\updatePcPerm{c_\adv}$\\
-\cline{2-2}
-$r_0$  &  $c_{\var{prp}}$ \\
-\cline{2-2}
-$r_1$ & $c_\adv$ \\
-\cline{2-2}
-$r_{\var{priv},1}$ & $w_{p,1}$\\
-\cline{2-2}
-$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
-\cline{2-2}
-& $0$ \\
-\cline{2-2}
-& $\vdots$ \\
-\cline{2-2}
-\end{tabular}
-\caption{Stack and register-file just after jump to $\adv$.}
-\label{fig:f2-stack-before-adv}
-\end{figure}
-Define 
-  \begin{align*}
-W = \;& [0 \mapsto \iota_{\malloc,0}] \\
-    & [1 \mapsto \iota^\sta(\perma,\ms_{f2})]\\ 
-    & [2 \mapsto \iota^\sta (\perma,\ms_\flag)]\\
-    & [3 \mapsto \iota^\sta (\perma,\ms_\link)]\\
-    & [4 \mapsto \iota^\nwl_{\start_\adv,\addrend_\adv} (\perma)]\\
-    & [5 \mapsto \iota^\sta (\temp,\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})] \\
-    & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk} (\temp)]
-  \end{align*}
-\lau{The interesting regions here are region 5 and 6.
-Region 5 governs our private part of the stack whereas region 6 governs the part of the stack we are about to release.
-I don't think it is strictly needed to make them $\temp$ for this example, but in general it is what we wish to do, so we can revoke them upon return.}
-Show
-\[
-  \npair{(\start_\adv,\addrend_\adv)}\in\readCond{}(\glob)(\iota^\pwl_{\start_\adv,\addrend_\adv})
-\]
-Which amounts to
-\[
-\iota^\nwl_{\start_\adv,\addrend_\adv} (\perma) \nsubeq[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv} (\perma)
-\]
-which is indeed the case. Use this to apply the FTLR to get 
-\begin{enumproof}
-\item $\npair{c_\adv} \in \stder(W)$
-\end{enumproof}
-Define
-\[
-  \ms  \defeq \hs_{f2} \uplus 
-        \hs_\flag \uplus                
-        \ms_\link \uplus 
-        \hs_\adv \uplus 
-        \ms_\malloc \uplus 
-        \ms_\stk'
-\]
-($m''$ without the frame) and show
-\begin{enumproof}[resume]
-\item $\heapSat[\ms]{n}{W}$ \lau{The argument here is a bit informal}
-  \begin{itemize}
-  \item For $\hs_{f2}$, $\hs_\flag$, and $\ms_\link$ the regions $\iota^\sta(\perma,\ms_{f2})$,$\iota^\sta (\perma,\ms_\flag)$, and $\iota^\sta (\perma,\ms_\link)$ are designed to respectively accept each of these memory segments.
-  \item $\hs_\adv$, showing this case is similar to item \ref{lem:f1-adv-mem-sat} in the proof of Lemma~\ref{lem:correctness-f1}.
-  \item $\ms_\malloc$, this case follows from the specification of malloc and the assumption that $\iota_{\malloc,0}$ models $\ms_\malloc$.    
-  \item $\ms_\stk'$, here we need to split the memory segment in two and for $n' < n$ argue
-    \begin{itemize}
-    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$:\\ This region matches only this memory segment.
-    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk'', \addrend_\stk]}} \in H_{\start_\stk'', \addrend_\stk}(1)(W)$:\\ Show $\forall a \in [\start_\stk'', \addrend_\stk] \ldotp \ms_\stk'|_{[\start_\stk'', \addrend_\stk]}(a) \in \stdvr(W)$, trivially true since these memory cells contain 0.
-    \end{itemize}
-  \end{itemize}
-\item $\npair{\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}} \in \stdrr(W)$
-  \begin{enumproof}
-    \item case $\npair{c_\stk''} \in \stdvr(W)$: Show:
-      \begin{enumproof}
-        \item $\npair{(\start_\stk'', \addrend_\stk)} \in \readCond{}(\local)(W)$ \\
-          This amounts to showing 
-\[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
-\]
-which is indeed the case.
-        \item $\npair{(\start_\stk'', \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$\\
-This amounts to showing 
-\[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
-\]
-which is indeed the case.
-        \item $\npair{(\rwlx,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$ \label{f2:rwlx-exec-case} \\
-We will show this case in some detail, the others are similar. \\
-Let $n' < n$, $W' \futurewk W$, and $a \in [\start_\stk'', \addrend_\stk]$ be given. Show
-\[
-  \npair[n']{((\rwlx,\local),\start_\stk'', \addrend_\stk,a)} \in \stder(W')
-\]
-If we can apply the FTLR at this point, then we are done. SFTS:
-\begin{enumproof}
-  \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \readCond{}(\local)(W')$\\
-this amounts to showing
-\[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsubeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
-\]
-which is indeed the case.
-  \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \writeCond{}(\local)(W')$ \\
-this amounts to showing
-\[
-  \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp) \nsupeq[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}(\temp)
-\]
-which is indeed the case.
-\end{enumproof}
-        \item $\npair{(\rwx,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-Similar to case~\ref{f2:rwlx-exec-case}
-        \item $\npair{(\exec,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-Similar to case~\ref{f2:rwlx-exec-case}
-      \end{enumproof}
-    \item case $\npair{c_\adv} \in \stdvr(W)$:\\
-     Follows from the FTLR.
-    \item case $\npair{c_{\var{prp}}} \in \stdvr(W)$:\\
-      Let $n' < n$ and $W' \future W$ be given and show
-      \[
-        \npair[n']{\updatePcPerm{c_{\var{prp}}}} \in \stder(W')
-      \]
-      to this en let $n'' \leq n'$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n'']{\reg'} \in \stdrr$ be given and show
-      \[
-        \npair[n'']{(\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms')} \in \observations(W')
-      \]
-      Let $i \leq n''$, $m''$ and $\ms_f$ be given and assume
-      \[
-        (\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms'\uplus\ms_f) \step[i]
-        (\halted,m'')
-      \]
-      (From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack).
-The restoration code remains unaltered, and so does the data on the stack, so the execution continues by poping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$.
-Continuing in $f2$, the restore code is poped off the stack which is the last of the $\mathtt{scall}$ instruction.
-Now the top word of the stack is poped to $r_1$, and thanks to the world satisfaction, we know that this word is 1.
-The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.)
-
-Use $\revokeTemp(W') \futurestr W'$. From the private-future memory satisfaction lemma (Lemma~\ref{lem:priv-mono-like}) get $\ms''$ and $\ms_f$ such that
-
-    \item case $r\not\in \{\pcreg,r_0,r_1,r_\stk\}$: $\reg_0(r) = 0$. Trivial.
-  \end{enumproof}
-\end{enumproof}
-to get
-
-\end{proof}
 
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^\pwl : \Views \fun \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq \lambda v \ldotp
-    \begin{cases}
-      (v,1,=,H^\pwl_{\start,\addrend}) & v \neq \revoked \\
-      \bot & \text{otherwise}
-    \end{cases} \\
-  &H^\pwl : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
-  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+  &\iota_{\start,\addrend}^{\var{pwl}} : \Views \fun \Regions\\
+  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda \_ \ldotp(\temp,(\start,\addrend),=,H^{\var{pwl}}) \\ \\
+  &H^{\var{pwl}} : \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
 \end{align*}
-\lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
 Revoking all temporary regions:
@@
     \begin{cases}
-      (v,1,=,H^\nwl_{start,\addrend}) & v \neq \revoked \\
+      (v,(\start,\addrend),=,H^{\var{nwl}}) & v \neq \revoked \\
       \bot & \text{otherwise}
@@
 \\ \\
-  &H^\nwl : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^{\var{nwl}} : \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
+  &H^{\var{nwl}} \; (\start,\addrend) \;\hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
@@
                         & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-                        & \;\qquad W(r)\nsubeq[n-1] \iota^\pwl_{\start',\addrend'}(W(r).v) \}
+                        & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
             \end{aligned}
@@
              & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \}
              \union \\
@@
              & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
+             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \land\\
              & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
@@
 
-\subsection{Useful regions}
-Static region used for parts of memory that should not change.
-\begin{align*}
-  \iota^\sta (v,\ms) &= (v,1,=,H^\sta\;\ms)\\
-  H^\sta \; \ms \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
-\end{align*}
-
-\begin{align*}
-  \iota^\cnst (v,n) &= (v,1,=,H^\cnst\;n)\\
-  H^\sta \; n \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \Rightarrow \forall a \in \dom(\ms)\ldotp \ms(a) = n \}
-\end{align*}
-
 \subsection{Lemmas}
@@
 \begin{lemma}[Private monotonicity-like property]
-\label{lem:priv-mono-like}
   \begin{align*}
@@
     & \qquad \Rightarrow \npair{\stdcap[(\perm,\glob)]} \in \stdvr(\wrev{W}) 
-  \end{align*}
-\end{lemma}
-
-\subsubsection{Standard regions}
-\begin{lemma}
-  \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubeq \iota^\pwl_{\start,\addrend}(v)
   \end{align*}

### 2016-11-30 local-capability-machine.tex
Public monotonicity of memory satisfaction does not hold (please read comment)

--- 2016-11-29 local-capability-machine.tex
@@
 
+\lau{I have skipped the following program in favor of the one in the next section. }
 ML-like program:
@@
 We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
+
+\begin{verbatim}
+f3: push 1
+    scall adv([],[])
+    pop r1
+    assert r1 1
+    push 2
+    scall adv([],[])
+3f: halt
+\end{verbatim}
 
@@
 \begin{lemma}[Public monotonicity]
+\lau{This is not true! If the new region is active and does not accept the empty memory, then it is no longer possible to make a satisfactory partition of the heap.}
   \begin{align*}

### 2016-12-01 local-capability-machine.tex
Typo fixes and some small remarks...

--- 2016-11-29 local-capability-machine.tex
@@
 \item Figure~\ref{fig:glob-hier} shows the ordering of $\local$ and $\glob$, i.e., the elements of $\Globals$.
-\item The ordering of $\Perms \times \Globals$ is point wise.
+\item The ordering of $\Perms \times \Globals$ is pointwise.
 \end{itemize}
@@
 But even if $w$ is just an integer, we still need the capability we write through to have some kind of write permission.}
+\dominique{I see, but maybe writeAllowed could take $w$ and $c$ as an argument, rather than just $\perm$? It is now weird that some conditions for writing are in $\writeAllowed{}$, but not all.}
+
+\dominique{factor out reoccuring conditions of the form $n_i = \rv_i$ or $n_i =
+  \memreg(\rv_i)$ into a function that evaluates an $\rv$ using a system state?
+  This would generalize nicely to the case where we would have additional
+  addressing modes.}
+
+\dominique{why doesn't store take an $\rv$?}
 
@@
 \section{Macros}
-In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
+In order to write readable example programs, we provide macros (macro-instructions) that can be implemented in terms of the instruction set given in the formalisation.
 \dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
@@
 }
+\lau{Agreed.}
+
 \dominique{what does ``fetch the capability ...'' mean?}
@@
   \label{lem:correctness-f2}
-  For all $n \in \nats$
   let
@@
     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
-    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\var{offset-link-flag},\mathtt{2f},\mathtt{f2}) \\
+    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-(\var{offset-link-flag}),\mathtt{2f},\mathtt{f2}) \\
     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
@@
     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
   \end{align*}
@@
 \]
-which is indeed the case. Use this to apply the FTLR to get 
+which is indeed the case\dominique{by some lemma?}. Use this to apply the FTLR to get 
 \begin{enumproof}
@@
     \begin{itemize}
-    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$:\\ This region matches only this memory segment.
+    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in
+      H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$ \dominique{apply to world?}:\\ This region matches only this memory segment.
     \item case $\npair[n']{\ms_\stk'|_{[\start_\stk'', \addrend_\stk]}} \in H_{\start_\stk'', \addrend_\stk}(1)(W)$:\\ Show $\forall a \in [\start_\stk'', \addrend_\stk] \ldotp \ms_\stk'|_{[\start_\stk'', \addrend_\stk]}(a) \in \stdvr(W)$, trivially true since these memory cells contain 0.
@@
 We will show this case in some detail, the others are similar. \\
+\dominique{Factor out a reusable lemma?}
 Let $n' < n$, $W' \futurewk W$, and $a \in [\start_\stk'', \addrend_\stk]$ be given. Show
@@
       \]
+      \dominique{part of a phrase missing here?}
       to this en let $n'' \leq n'$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n'']{\reg'} \in \stdrr$ be given and show
@@
 The restoration code remains unaltered, and so does the data on the stack, so the execution continues by popping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$.
-Continuing in $f2$, the restore code is popped off the stack which is the last of the $\mathtt{scall}$ instruction.
+Continuing in $f2$, the restore code is popped off the stack which is the last of the $\mathtt{scall}$ macro-instruction.
 Now the top word of the stack is popped to $r_1$, and thanks to the world satisfaction, we know that this word is 1.
 The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.
+      \dominique{Maybe you can factor out some sort of reusable antireduction lemma
+        for this sort of argument?}
 
@@
 \]
-Now using $i$, the number of steps we need to halt, $\ms_f$ as the frame, $m'$ as the resulting memory and the assumption that the execution terminates successfuly gives $W' \futurestr W$, $\ms'$, and $\ms_r$ such that
+Now using $i$, the number of steps we need to halt, $\ms_f$ as the frame, $m'$ as the resulting memory and the assumption that the execution terminates successfully gives $W' \futurestr W$, $\ms'$, and $\ms_r$ such that
 \begin{itemize}

### 2016-12-01 local-capability-machine.tex
Many things happened. Sorry, Dominique

--- 2016-11-30 local-capability-machine.tex
@@
 \newcommand{\cnst}{\var{cnst}}
+\newcommand{\olf}{\var{offsetLinkFlag}}
+\newcommand{\prp}{\var{prp}}
 
@@
 \newcommand{\erase}[2]{\floor*{#1}_{\{#2\}}}
+\newcommand{\activeReg}[1]{\plainfun{active}{#1}}
 
@@
 
+\newcommand{\localityReg}{\var{localityReg}}
 \newcommand{\localReg}{\var{localReg}}
@@
 \[
-    \var{offset-link-flag} \defeq 2
+    \olf \defeq 2
 \]
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
-    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\var{offset-link-flag},\mathtt{1f},\mathtt{f1}) \\
-    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\var{offset-link-flag}) \\
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
     m & \defeq \hs_{f1} \uplus 
@@
 \begin{align*}
-  &\dom(\hs_{f1}) = [\mathtt{f1}-\var{offset-link-flag},\mathtt{1f}] \\
+  &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
   &\dom(\hs_\flag) = [\flag,\flag] \\
@@
 \item $\ms_{\var{ar}}'$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$ (the return address in f1), and activation code.
-\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+\var{offset-link-flag})$ where $b+\var{offset-link-flag}$ is the first address of the activation code.
+\item $c_{\var{ar}}' = ((\entry,\local)b,e,b+\olf)$ where $b+\olf$ is the first address of the activation code.
 \end{itemize}
@@
 
-\begin{lemma}
+\begin{lemma}[Correctness lemma for \texttt{f2}]
   \label{lem:correctness-f2}
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\var{offset-link-flag}) \\
-    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\var{offset-link-flag},\mathtt{2f},\mathtt{f2}) \\
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
@@
   \begin{align*}
-    &\dom(\hs_{f2}) = [\mathtt{f2}-\var{offset-link-flag},\mathtt{2f}] \\
+    &\dom(\hs_{f2}) = [\mathtt{f2}-\olf,\mathtt{2f}] \\
     &\dom(\hs_\flag) = [\flag,\flag] \\
@@
 $r_1$ & $c_\adv$ \\
-\cline{2-2}
-$r_{\var{priv},1}$ & $w_{p,1}$\\
 \cline{2-2}
@@
      Follows from the FTLR.
-    \item case $\npair{c_{\var{prp}}} \in \stdvr(W)$:\\
+    \item case $\npair{c_{\var{prp}}} \in \stdvr(W)$: \label{f2:c-prp}\\
       Let $n' < n$ and $W' \future W$ be given and show
@@
 \end{verbatim}
+
+\begin{lemma}[Correctness lemma for \texttt{f3}]
+    \label{lem:correctness-f3}
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f3} & \defeq ((\rwx,\glob),\mathtt{f3}-\olf,\mathtt{3f},\mathtt{f3}) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
+    \reg & \in \Regs \\
+    m & \defeq \hs_{f3} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus 
+        \ms_{\malloc} \uplus 
+        \ms_{\var{stk}} \uplus
+        \ms_{\var{frame}} 
+  \end{align*}
+  where 
+  \begin{align*}
+    &\dom(\hs_{f3}) = [\mathtt{f3}-\olf,\mathtt{3f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link+1]\\
+    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item The first address of $\hs_{f3}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f3}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f3}$ contains the code of $f3$.
+  \item $\ms_\flag = [flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of code.
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
+This proof goes like the proof of Lemma~\ref{lem:correctness-f3} except for item \ref{f2:c-prp}
+\begin{enumproof}[start=3]
+  \item $\npair{\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}} \in \stdrr(W)$ \\
+    $\dots$
+  \begin{enumproof}[start=3]
+  \item $\npair{c_{\var{prp}}} \in \stdvr(W)$ \\
+    To this end let $n' < n$ and $W' \futurewk W$ be given and show
+    \[
+      \npair[n']{\updatePcPerm(c_\prp)} \in \stder(W')
+    \]
+    to this end let 
+    \begin{enumproof}
+      \item $n'' \leq n'$
+      \item $\npair[n'']{\reg} \in \stdrr(W')$ \label{f3:mem-sat-mid}
+      \item $\heapSat[\ms]{n''}{W'}$
+    \end{enumproof}
+    be given and show
+    \[
+      \npair[n'']{(\reg\update{\pcreg}{\updatePcPerm(c_\prp)},\ms)}
+    \]
+    to this end let $i \leq n''$, $\ms_f$, and $m''$ be given and assume
+    \begin{enumproof}[resume]
+      \item $(\reg\update{\pcreg}{c_\prp},\ms \uplus \ms_f) \step[i] (\halted,m'')$ \label{f3:mid-exec}
+    \end{enumproof}
+    now show
+    \[
+      \exists \ms_R,\ms', W'' \futurestr W' \ldotp m'' = \ms' \uplus \ms_R \uplus \ms_f \land \heapSat[\ms']{n''-i}{W''}
+    \]
+    At this point use Lemma~\ref{lem:priv-mono-like2} with \ref{f3:mem-sat-mid} to get $\ms_p$, $\ms_r$, and $R$ that such that $\ms = \ms_p \uplus \ms_r$ and $\heapSat[\ms_p]{n''}{W'}$.
+We can further conclude that $\ms |_{[\start_\stk,\addrend_\stk]} \subseteq \ms_r$ \lau{The needs to be spelled out - preferably in an easy to use lemma.
+The idea is that $R$ partitions $\ms_r$ into memory segments that each are accepted by a temporary region.
+The stack is governed by a temporary region, so it must be in $\ms_r$.}.
+
+ Now consider the first steps of execution \ref{f3:mid-exec}.
+From \ref{f3:mem-sat-mid} we have that the private part of the stack and $f3$ is unchanged, so the recovery code gets to run which restores the stack capability and program counter, and pops the return code off the stack (there are no ``private'' registers).
+Next, the top word is popped off the stack.
+Again due to memory satisfaction, we know that this word is the integer 1 and therefore the assertion succeeds.
+Execution continues by pushing 2 to the stack.
+Now due the $\mathtt{scall}$ up to the point just after the jump to $\adv$.
+Figure~\ref{fig:f3-stack-before-snd-adv} shows the stack and register file at this point.
+The $\ms_p$ part of $\ms$ was not changed.
+For convenience, name the stack part of memory at this point $\ms'_\stk$.
+
+Let $W^{(3)} = \revokeTemp{W''}\update{i_1}{\iota^\sta(\temp,\ms|_{[\start_\stk, \start_\stk''-1]})}\update{i_2}{\iota^\sta(\temp,\ms|_{[\start_\stk'',\addrend_\stk]})}$
+
+Show
+
+\begin{figure}
+  \centering
+  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
+\multicolumn{2}{l}{Stack} \\
+\cline{2-2}
+   &         \\
+   & $\vdots$ \\
+\cline{2-2}
+$\start_\stk'':$   & $0$\\
+\cline{2-2}
+$c_\stk',c_\stk'' \rightarrow$   & $c_\stk'$ \\
+\cline{2-2}
+   & $c_{f2}'$ \\
+\cline{2-2}
+$c_{\var{prp}} \rightarrow$   & Restore code \\
+   & $\vdots$ \\
+\cline{2-2}
+   & $2$\\
+\cline{2-2}
+\end{tabular}
+\hspace{1cm}
+\begin{tabular}{r |c |}
+\multicolumn{2}{r}{Register file} \\
+\cline{2-2}
+$\pcreg$ & $\updatePcPerm{c_\adv}$\\
+\cline{2-2}
+$r_0$  &  $c_{\var{prp}}$ \\
+\cline{2-2}
+$r_1$ & $c_\adv$ \\
+\cline{2-2}
+$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
+\cline{2-2}
+& $0$ \\
+\cline{2-2}
+& $\vdots$ \\
+\cline{2-2}
+\end{tabular}
+\caption{Stack and register-file just after second jump to $\adv$.}
+\label{fig:f3-stack-before-snd-adv}
+\end{figure}
+  \end{enumproof}
+\end{enumproof}
+\end{proof}
 
@@
 
+Define the function $\active{}$ as follows
+\begin{align*}
+  \activeReg{} & : \Worlds \fun 2^\RegionNames \\
+  \activeReg{W} & \defeq \dom(\erase{W}{\perma,\temp})
+\end{align*}
+
 Memory segment satisfaction:
@@
   \left\{\begin{aligned}
-    &\exists R : \dom(\erase{W}{\perma,\temp}) \rightarrow \HeapSegments \ldotp \\
-    &\quad \hs = \uplus_{r\in\dom(\erase{W}{\perma,\temp})}R(r) \land \\
-    &\quad \forall r \in \dom(\erase{W}{\perma,\temp}) \ldotp\\
+    &\exists R : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
+    &\quad \hs = \biguplus_{r\in\activeReg{W}}R(r) \land \\
+    &\quad \forall r \in \activeReg{W} \ldotp\\
     &\qquad \forall n'<n \ldotp \\
@@
 For convenience define
-\begin{align*}
-  \localReg(W) & \defeq \dom(\erase{W}{\perm,\temp}) \\
-  \globalReg(W) & \defeq \dom(\erase{W}{\perm})
-\end{align*}
-$\localReg$ are the regions that local capabilities may govern - that is permanent and temporary regions.
-$\globalReg$ are the regions that global capabilities may govern - that is permanent regions.
+\[
+  \localityReg(\gl,W) \defeq 
+  \begin{cases}
+    \dom(\erase{W}{\perm,\temp}) & \text{if } \gl = \local \\
+    \dom(\erase{W}{\perm}) & \text{if } \gl = \glob
+  \end{cases}
+\]
+$\localityReg(\local,W)$ are the regions that local capabilities may govern - that is permanent and temporary regions.
+$\localityReg(\glob,W)$ are the regions that global capabilities may govern - that is permanent regions.
 Now define the following function
-\[
-  \var{localityReg}(\gl,W) \defeq 
-  \begin{cases}
-    \localReg(W) & \text{if } \gl = \local \\
-    \globalReg(W) & \text{if } \gl = \glob
-  \end{cases}
-\]
 
@@
 \begin{align*}
-    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+    & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
     & \writeCond{}(\iota,\gl)(W) =  \\
@@
               & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \}
+            \end{aligned}
+\end{align*}
+
+\begin{align*}
+    & \readCond{} : \Globals \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+    & \readCond{}(\gl)(W) =  \\
+    & \quad \begin{aligned}[t]
+              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+              & \;\qquad W(r)\nsubsim[n-1] \iota_{\start',\addrend'}^\pwl \}
             \end{aligned}
@@
 
-\begin{lemma}[Private monotonicity-like property]
+\begin{lemma}[Revoke temporary memory satisfaction]
 \label{lem:priv-mono-like}
@@
     &      \forall \hs, n, W, W' \ldotp \\
-    & \quad  \heapSat[\hs]{n}{W} \land W' \futurewk \wrev{W} \futurestr W \\
-    & \qquad \Rightarrow \exists \hs', \hs_f \ldotp \hs = \hs' \uplus \hs_f \land \heapSat[\hs']{n}{W'}
+    & \quad  \heapSat[\hs]{n}{W} \Rightarrow\\
+    & \qquad \exists \hs', \hs_r \ldotp \\
+    & \qquad \quad \hs = \hs' \uplus \hs_r \land \heapSat[\hs']{n}{\wrev{W}}
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Revoke temporary memory satisfaction 2]
+\label{lem:priv-mono-like2}
+  \begin{align*}
+    & \forall \ms, n, W \ldotp \heapSat[\ms]{n}{W} \Rightarrow \\
+    & \quad \exists \ms', \ms_r, R : \dom(\erase{W}{\temp}) \fun \MemSegments \ldotp \\
+    & \qquad \ms = \ms' \uplus \ms_r \land \\
+    & \qquad \heapSat[\ms']{n}{\wrev{W}} \land \\
+    & \qquad \forall r \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).\mathrm{H}(W(r).\mathrm(s))(W)\\
+    & \qquad \ms_r = \biguplus_{r \in \dom(R)} R(r)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}[Disjoint memory satisfaction]
+\label{lem:disj-mem-sat}
+  \begin{align*}
+    & \forall n \ldotp \forall \ms, \ms' \ldotp \forall W, W' \ldotp\\
+    & \quad \heapSat[\ms \uplus \ms' ]{n}{W \uplus W'}
   \end{align*}

### 2016-12-01 local-capability-machine.tex
Addressed some of Dominique's comments.

@@
 \begin{itemize}
-\item The first address of $\hs_{f1}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f1}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f1}$ contains the code of $f1$.
-\dominique{why not write this in symbols?}
+\item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
 \item $\ms_\flag = [flag \mapsto 0]$
@@
   \begin{itemize}
-  \item The first address of $\hs_{f2}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f2}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f2}$ contains the code of $f2$.
+  \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
   \item $\ms_\flag = [flag \mapsto 0]$
@@
 \]
-which is indeed the case\dominique{by some lemma?}. Use this to apply the FTLR to get 
+which is indeed the case by Lemma~\ref{lem:nwlp-subset-pwl}. Use this to apply the FTLR to get 
 \begin{enumproof}
@@
     \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in
-      H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})$ \dominique{apply to world?}:\\ This region matches only this memory segment.
+      H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})(W)$ \dominique{apply to world?}:\\ This region matches only this memory segment.
     \item case $\npair[n']{\ms_\stk'|_{[\start_\stk'', \addrend_\stk]}} \in H_{\start_\stk'', \addrend_\stk}(1)(W)$:\\ Show $\forall a \in [\start_\stk'', \addrend_\stk] \ldotp \ms_\stk'|_{[\start_\stk'', \addrend_\stk]}(a) \in \stdvr(W)$, trivially true since these memory cells contain 0.
@@
       \[
-        \npair[n']{\updatePcPerm{c_{\var{prp}}}} \in \stder(W')
-      \]
-      \dominique{part of a phrase missing here?}
-      to this en let $n'' \leq n'$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n'']{\reg'} \in \stdrr$ be given and show
+        \npair[n']{\updatePcPerm{c_{\var{prp}}}} \in \stder(W').
+      \]
+      To this end let $n'' \leq n'$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n'']{\reg'} \in \stdrr$ be given and show
       \[
@@
   \begin{itemize}
-  \item The first address of $\hs_{f3}$ contains a global read-only capability for $\hs_\link$, the second address of $\hs_{f3}$ contains a global read-write capability to $\hs_\flag$, the rest of $\hs_{f3}$ contains the code of $f3$.
+  \item $\ms_{f3}(\mathtt{f3}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f3}(\mathtt{f3}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f3}$ contains the code of $f3$.
   \item $\ms_\flag = [flag \mapsto 0]$
@@
 \begin{lemma}
+\label{lem:nwl-subset-pwl}
   \begin{align*}
@@
     & \quad \iota^\nwl_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
+  \end{align*}
+\end{lemma}
+\begin{lemma}
+\label{lem:nwlp-subset-pwl}
+  \begin{align*}
+    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
+    & \quad \iota^{\nwl,p}_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
   \end{align*}

### 2016-12-02 local-capability-machine.tex
elaborated a comment
f3 proof done

@@
 \newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
+\newcommand{\memSat}[3][n]{\heapSat[#2]{#1}{#3}}
+\newcommand{\memSatPar}[4][n]{\heapSat[#2]{#1 , #4}{#3}}
 
@@
   \begin{enumproof}
-    \item case $\npair{c_\stk''} \in \stdvr(W)$: Show:
+    \item case $\npair{c_\stk''} \in \stdvr(W)$: Show: \label{f2:stack-in-vr}
       \begin{enumproof}
@@
       \item $n'' \leq n'$
-      \item $\npair[n'']{\reg} \in \stdrr(W')$ \label{f3:mem-sat-mid}
-      \item $\heapSat[\ms]{n''}{W'}$
+      \item $\npair[n'']{\reg} \in \stdrr(W')$
+      \item $\heapSat[\ms]{n''}{W'}$ \label{f3:mem-sat-mid}
       \end{enumproof}
@@
       now show
-    \[
-      \exists \ms_R,\ms', W'' \futurestr W' \ldotp m'' = \ms' \uplus \ms_R \uplus \ms_f \land \heapSat[\ms']{n''-i}{W''}
-    \]
-    At this point use Lemma~\ref{lem:priv-mono-like2} with \ref{f3:mem-sat-mid} to get $\ms_p$, $\ms_r$, and $R$ that such that $\ms = \ms_p \uplus \ms_r$ and $\heapSat[\ms_p]{n''}{W'}$.
-We can further conclude that $\ms |_{[\start_\stk,\addrend_\stk]} \subseteq \ms_r$ \lau{The needs to be spelled out - preferably in an easy to use lemma.
-The idea is that $R$ partitions $\ms_r$ into memory segments that each are accepted by a temporary region.
-The stack is governed by a temporary region, so it must be in $\ms_r$.}.
-
- Now consider the first steps of execution \ref{f3:mid-exec}.
+
+      \begin{enumproof}
+        \item $\exists \ms_R,\ms', W'' \futurestr W' \ldotp m'' = \ms' \uplus \ms_R \uplus \ms_f \land \heapSat[\ms']{n''-i}{W''}$ \label{f3:proof-burden}.
+      \end{enumproof}
+      From \ref{f3:mem-sat-mid}, get $P$ such that 
+      \begin{enumproof}[resume]
+      \item $\memSatPar[n'']{\ms}{W'}{P}$ \label{f3:mem-sat-par-mid}
+      \end{enumproof}
+      Use this with Lemma~\ref{lem:priv-mono-like2} to get $\ms_r$ and $\ms'$ where
+      \begin{enumproof}[resume]
+      \item $\ms = \ms' \uplus \ms_r$
+      \item $\memSatPar{\ms'}{\revokeTemp{W'}}{P|_{\dom(\erase{W'}{\perma})}}$ \label{f3:perm-mem-sat}
+      \item $\ms_r = \biguplus_{r \in \erase{W'}{\temp}} P(r)$ \label{f3:ms-revoked}
+      \item $\ms' = \biguplus_{r \in \erase{W'}{\perma}} P(r)$
+      \end{enumproof}
+      Let
+      \begin{itemize}
+      \item $\ms_{\stk,\var{priv}} = \ms |_{\start_\stk,\base_\stk''-1}$
+      \item $\ms_{\stk,\var{rest}} = \ms |_{\base_\stk'',\addrend_\stk}$
+      \end{itemize}
+      We know $W'(6)=\iota^\pwl_{\base_\stk'',\addrend_\stk}$ (using the definition of $W$ from the proof of Lemma~\ref{lem:correctness-f2}) which is a $\temp$ region.
+We futher know that $P(6) = \ms_{\var{?}}$ for some memory segment.
+By \ref{f3:mem-sat-par-mid} we know that $\npair[n'']{\ms_?} \in H^\pwl_{\base_\stk'',\addrend_\stk} \; 1 \; W'$ and thus $\ms_?
+= \ms_{\stk,\var{rest}}$.
+Sine $W'(6)$ is a $\temp$ region, we can from \ref{f3:ms-revoked} conlude that
+      \begin{enumproof}[resume]
+      \item $\ms_{\stk,\var{rest}} \subseteq \ms_r$
+      \end{enumproof}
+      In a similar fashion, get
+      \begin{enumproof}[resume]
+      \item $\ms_{\stk,\var{priv}} \subseteq \ms_r$
+      \end{enumproof}
+
+      Now consider what happens in execution \ref{f3:mid-exec}.
 From \ref{f3:mem-sat-mid} we have that the private part of the stack and $f3$ is unchanged, so the recovery code gets to run which restores the stack capability and program counter, and pops the return code off the stack (there are no ``private'' registers).
@@
 Again due to memory satisfaction, we know that this word is the integer 1 and therefore the assertion succeeds.
+Notice that up to this point, no change has been done to memory.
 Execution continues by pushing 2 to the stack.
-Now due the $\mathtt{scall}$ up to the point just after the jump to $\adv$.
+Now consider the execution of $\mathtt{scall}$ up to the point just after the jump to $\adv$.
+Restore code is pushed to the stack, return and stack capabilities are pushed to the stack, and the rest of the stack is cleared.
 Figure~\ref{fig:f3-stack-before-snd-adv} shows the stack and register file at this point.
-The $\ms_p$ part of $\ms$ was not changed.
-For convenience, name the stack part of memory at this point $\ms'_\stk$.
-
-Let $W^{(3)} = \revokeTemp{W''}\update{i_1}{\iota^\sta(\temp,\ms|_{[\start_\stk, \start_\stk''-1]})}\update{i_2}{\iota^\sta(\temp,\ms|_{[\start_\stk'',\addrend_\stk]})}$
-
+Let $\ms_\stk'$ be the stack at this point.
+
+      For convenience define let $i_1 \neq i_2$ and $i_1,i_2 \not\in \dom(\revokeTemp(W'))$ and define
+      \[
+        W^{(3)} = \revokeTemp{W'}\update{i_1}{\iota^\pwl_{\base_\stk'',\addrend_\stk}}\update{i_2}{\iota^\sta (\temp,\ms_\stk' |_{\start_\stk,\base_\stk''-1})}
+      \]
+      We now want to use the FTLR on the $\updatePcPerm{c_\adv}$ capability which is possible as we have 
+      \[
+        \npair[n'']{\start_\adv,\addrend_\adv} \in \readCond(\glob)(W^{(3)})
+      \]
+      This follows from Lemma~\ref{lem:nwlp-subset-pwl}.
+
+      We have 
+      \[
+        \npair[n''] \in \stder(W^{(3)})
+      \]
       Show
-
+      \begin{enumproof}[resume]
+      \item $\memSat[n'']{\ms' \uplus \ms_\stk'}{W^{(3)}}$ \\
+        show:
+        \begin{itemize}
+        \item $\memSat[n'']{\ms_\stk'}{[i_1 \mapsto \iota^\pwl_{\base_\stk'',\addrend_\stk}][i_2\mapsto \iota^\sta (\temp,\ms_\stk' |_{\start_\stk,\base_\stk''-1})]}$ 
+        \end{itemize}
+        The private part of the stack is exactly the memory of the that is accepted by region $i_2$ and the rest of the stack contains all 0's, so it is accepted by region $i_1$.
+
+        Using the above and \label{f3:perm-mem-sat} by Lemma~\ref{lem:disj-mem-sat} get
+        \begin{enumproof}
+        \item $\memSat[n'']{\ms' \uplus \ms_\stk'}{\revokeTemp{W'}\update{i_1}{\iota^\pwl_{\base_\stk'',\addrend_\stk}}\update{i_2}{\iota^\sta (\temp,\ms_\stk' |_{\start_\stk,\base_\stk''-1})}}$ 
+        \end{enumproof}
+        Notice that this is only possible, because we argued that the stack part was in $\ms_r$, so it is disjoint from $\ms'$. The lower step comes from downwards closure.
+      \item $\npair[n'']{\reg_0\update{r_0}{c_{\var{prp}}'}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk'}} \in \stdrr(W^{(3)})$ \\
+        Amounts to showing:
+        \begin{enumproof}
+          \item $\npair[n'']{c_\adv} \in \stdvr(W^{(3)})$\\
+            Follows from FTLR.
+          \item $\npair[n'']{c_\stk'} \in \stdvr(W^{(3)})$\\
+            Like case \ref{f2:stack-in-vr} in the proof of Lemma~\ref{lem:correctness-f2}.
+          \item $\npair[n'']{c_{\var{prp}}} \in \stdvr(W^{(3)})$\\
+            Let $n^{(3)} < n''$ and $W^{(4)} \futurewk W^{(3)}$ be given. Further led $\ms''$and $\reg'$ be given such that:
+            \begin{itemize}
+            \item $\npair[n^{(3)}]{\reg'} \in \stdrr(W^{(4)})$
+            \item $\memSat[n^{(3)}]{\ms''}{W^{(4)}}$
+            \end{itemize}
+            and show
+            \begin{enumproof}
+              \item $\npair[n^{(3)}]{(\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms'')} \in \observations(W^{(4)})$\\
+                To this end let $\ms_f''$, $k \leq n^{(3)}$, $m^{(3)}$ be given and assume
+                \[
+                  (\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms' \uplus \ms_f'') \step[k] (\halted,m^{(3)})
+                \]
+                From $\memSat[n^{(3)}]{\ms'}{W^{(4)}}$ we know that the $\ms_f$ and $\ms_\stk' |_{\base_\stk,\base_\stk''-1}$ is part of $\ms''$, so upon return, the restore code is run and subsequently, the $\mathtt{halt}$ instruction in $f3$ is executed.
+No changes have been made to the memory, so $\memSat[n^{(3)}]{\ms'}{W^{(4)}}$ provides the desired result (picking the ``revoked'' part to be empty and $W^{(4)}$ to be the private future world).
+            \end{enumproof}
+        \end{enumproof}
+      \end{enumproof}
+      to get
+      \[
+        \npair[n'']{(\reg_0\update{r_0}{c_{\var{prp}}'}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk'}\update{\pcreg}{\updatePcPerm{c_\adv}},\ms' \uplus \ms_\stk')} \in \observations(W^{(3)})
+      \]
+      By assumption $\ref{f3:mid-exec}$ the execution terminates after $i-j$ steps in memory $m''$ where $j$ is the amount of steps we described previously.
+The above gives us $W^{(4)} \futurestr W^{(3)}$, $\ms''$, and $\ms_R'$ such that
+      \begin{itemize}
+      \item $m'' = \ms'' \uplus \ms_R' \uplus \ms_f$
+      \item $\memSat[n''-(i-j))]{\ms''}{W^{(4)}}$
+      \end{itemize}
+      Using downwards closure of heap satisfaction, this is exactly what we needed to show \ref{f3:proof-burden}
       \begin{figure}
@@
           \cline{2-2}
-   & $c_{f2}'$ \\
+       & $c_{f3}''$ \\
           \cline{2-2}
@@
           \cline{2-2}
-$r_0$  &  $c_{\var{prp}}$ \\
+          $r_0$  &  $c_{\var{prp}}'$ \\
           \cline{2-2}
@@
   \left\{\begin{aligned}
-    &\exists R : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
-    &\quad \hs = \biguplus_{r\in\activeReg{W}}R(r) \land \\
-    &\quad \forall r \in \activeReg{W} \ldotp\\
-    &\qquad \forall n'<n \ldotp \\
-    &\qquad \quad \exists H,s \ldotp\\
-    &\qquad \qquad W(r) = (\_,s,\_,H) \land \\
-    &\qquad \qquad \npair[n']{R(r)} \in H(s)(\xi^{-1}(W))\\
+    &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
+    &\quad \memSatPar{\ms}{W}{P}
+  \end{aligned}\right.
+\end{align*}
+
+\begin{align*}
+  &\memSatPar{\ms}{W}{P}
+  \text{ iff }
+  \left\{\begin{aligned}
+    &\hs = \biguplus_{r\in\activeReg{W}}P(r) \land \\
+    &\forall r \in \activeReg{W} \ldotp\\
+    &\quad  \forall n'<n \ldotp \\
+    &\qquad \exists H,s \ldotp\\
+    &\qquad \quad W(r) = (\_,s,\_,H) \land \\
+    &\qquad \quad \npair[n']{P(r)} \in H(s)(\xi^{-1}(W))\\
   \end{aligned}\right.
@@
 For instance, it was not possible to have a global read capability for a region governed by a permanent standard region.
+The issue was that $\iota^\pwl$ is always $\temp$ which gave some issues in the read and write conditions when they used $\iota^\pwl$.
+Another option was to be more explicit, but we opted for this more elegant(?) solution.
 }
@@
   \begin{align*}
-    & \forall \ms, n, W \ldotp \heapSat[\ms]{n}{W} \Rightarrow \\
-    & \quad \exists \ms', \ms_r, R : \dom(\erase{W}{\temp}) \fun \MemSegments \ldotp \\
-    & \qquad \ms = \ms' \uplus \ms_r \land \\
-    & \qquad \heapSat[\ms']{n}{\wrev{W}} \land \\
-    & \qquad \forall r \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).\mathrm{H}(W(r).\mathrm(s))(W)\\
-    & \qquad \ms_r = \biguplus_{r \in \dom(R)} R(r)
+    & \forall \ms, n, W, R : \activeReg(W) \fun \MemSegments  \ldotp \\
+    & \quad \memSatPar{\ms}{W}{P} \Rightarrow \\
+    & \qquad \exists \ms', \ms_r\ldotp \\
+    & \qquad \quad \ms = \ms' \uplus \ms_r \land \\
+    & \qquad \quad \memSatPar{\ms'}{\wrev{W}}{P|_{\dom(\erase{W}{\perma})}} \land \\
+    & \qquad \quad \ms_r = \biguplus_{r \in \erase{W}{\temp}} P(r) \land \\
+    & \qquad \quad \ms' = \biguplus_{r \in \erase{W}{\perma}} P(r)
   \end{align*}
@@
     & \forall n \ldotp \forall \ms, \ms' \ldotp \forall W, W' \ldotp\\
-    & \quad \heapSat[\ms \uplus \ms' ]{n}{W \uplus W'}
+    & \quad \memSat{\ms}{W} \land \memSat{\ms'}{W'} \Rightarrow \\
+    & \qquad \heapSat[\ms \uplus \ms' ]{n}{W \uplus W'}
   \end{align*}

### 2016-12-02 local-capability-machine.tex
Added reqglob macro that requires capability in register to be global
Added reqperm

@@
 \newcommand{\encodeLoc}{\mathit{encodeLoc}{}}
-\newcommand{\decodePermPair}{\plainfun{decodePerm}}
+\newcommand{\decodePermPair}{\plainfun{decodePermPair}}
+\newcommand{\decodePerm}[1]{\plainfun{decodePerm}{#1}}
 \newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}
@@
 \end{description}
+\item[\texttt{reqglob $r$}] Tests if register $r$ contains a $\glob$ capability. If not fail, otherwise continue execution.
+\item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with $\decodePerm{n}$. If not fail, otherwise continue execution.
 \item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 
@@
  matters simple, we opt not to do so.
+\item \texttt{reqperm} can be used to test whether something can pass as a stack.
+\item \texttt{reqglob} can be used to test whether a callback is admissable in the presence of a stack.
 \end{itemize}

### 2016-12-05 local-capability-machine.tex
Added create closure instruction

@@
 \end{description}
-\item[\texttt{reqglob $r$}] Tests if register $r$ contains a $\glob$ capability. If not fail, otherwise continue execution.
-\item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with $\decodePerm{n}$. If not fail, otherwise continue execution.
-\item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 
 \end{description}
@@
 If any callbacks are provided, then they need to be global.
+\end{itemize}
+\begin{description}
+\item[\texttt{crtcls $[(x_1,r_1),\dots(x_n,r_n)]$ $r_{\var{code}}$}] \forcenewline
+$[(x_1,r_1),\dots(x_n,r_n)]$ is a list of variable bindings.
+If an instruction refers to a variable, then it will assume that an environment is available in a designated register (say $r_{\var{env}}$).
+The register $r_{\var{code}}$ should contain a capability governs the code of the closure and that is executable when jumped to.
+  \begin{description}
+  \item[Allocate memory for variable environment]
+  \item[Store register contents to environment]
+  \item[Allocate memory for record with environment capability, code capability, and acitvation code]
+  \item[Store capabilities and activation code to record]
+  \item[Restrict the capability for the ``clojure pair'' to an enter capability]
+  \item[Activation code:] \forcenewline
+    \begin{itemize}
+    \item Load the environment capability to a designated register
+    \item Load the code capability.
+    \item Jump to the code.
+    \end{itemize}
+  \end{description}
+  A more detailed description of each step:
+  \begin{description}
+  \item[Allocate memory for variable environment] Have malloc allocate a piece of memory of size $n$ (the size of the variable environment). 
+  \item[Store register contents to environment] Store the contents of each of the registers $r_1,\dots,r_n$ to the newly allocated memory
+  \item[Allocate memory for record with environment capability, code capability, and acitvation code] Allocate a new piece of memory with room for a capability for the environment
+  \item[Store capabilities and activation code to record] Store the environment capability and code capability in the record followed by the activation code. 
+  \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the acitvation code and restrict it to an enter capability.
+  \item[Activation code:] \forcenewline
+    \begin{itemize}
+    \item Load the environment capability to a designated register.
+    \item Load the code capability.
+    \item Jump to the code.
+    \end{itemize}
+  \end{description}
+\item[\texttt{load $r$ $x$}] Assumes environment capability available in register $r_{\var{env}}$. Loads the at the index associated with $x$ in the list. Loads from this capability into $r$.
+\item[\texttt{store $x$ $r$}] Assumes environment capability available in register $r_{\var{env}}$.
+Loads the at the index associated with $x$ in the list.
+Stores the contents of register $r$ through this capability.
+\item[\texttt{reqglob $r$}] Tests if register $r$ contains a $\glob$ capability. If not fail, otherwise continue execution.
+\item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with $\decodePerm{n}$. If not fail, otherwise continue execution.
+\item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 
+\end{description}
+Note:
+\begin{itemize}
 \item In a real setting due to a limited number of registers, some of the arguments might be spilled to the stack. It would be possible to do something similar here, but to keep
@@
 \item \texttt{reqglob} can be used to test whether a callback is admissable in the presence of a stack.
+\item The code of a closure will often be found in conjunction with the code that creates it.
 \end{itemize}

### 2016-12-05 local-capability-machine.tex
first version of assembly awkward example

@@
 \usepackage{mathpartir}
+\usepackage{listings}
+\lstset{
+  basicstyle=\ttfamily,
+  columns=fullflexible,
+  keepspaces=true,
+  mathescape
+}
 
@@
 
-The stack resides entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing.
+The stack resides entirely in memory. There is no separation between the memory and the stack, so when we talk about the stack it is as a conceptual thing. %
 
@@
   \end{description}
-\item[\texttt{load $r$ $x$}] Assumes environment capability available in register $r_{\var{env}}$. Loads the at the index associated with $x$ in the list. Loads from this capability into $r$.
+\item[\texttt{load $r$ $x$}] Assumes environment capability available in register $r_{\var{env}}$.
+Loads the word at the index associated with $x$ in the environment list.
+Loads from this capability into $r$.
 \item[\texttt{store $x$ $r$}] Assumes environment capability available in register $r_{\var{env}}$.
-Loads the at the index associated with $x$ in the list.
+Loads the word at the index associated with $x$ in the environment list.
 Stores the contents of register $r$ through this capability.
@@
 \end{verbatim}
+\begin{lstlisting}
+g2:  malloc $r_2$ 1
+     store $r_2$ 0
+     move $\pcreg$ $r_3$
+     lea $r_3$  $\dots$
+     crtcls $[(x, r_2)]$ $r_3$
+     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
+2g:  jmp $r_0$
+
+f4:  reqglob $r1$
+     reqperm $r_\stk$ $\encodePerm(\rwlx)$
+     store $x$ 0
+     scall $r_1$($[]$,$[r_0,r_1]$)
+     store $x$ 1
+     scall $r_1$($[]$,$[r_0]$)
+     load $r_1$ x
+     assert $r_1$ 1
+     cclear $\RegName \setminus \{r_0,\pcreg\}$
+4f:  jmp $r_0$
+\end{lstlisting}
 

### 2016-12-06 local-capability-machine.tex
g1 lemma

--- 2016-12-05 local-capability-machine.tex
@@
   \item[Allocate memory for variable environment] Have malloc allocate a piece of memory of size $n$ (the size of the variable environment). 
-  \item[Store register contents to environment] Store the contents of each of the registers $r_1,\dots,r_n$ to the newly allocated memory
-  \item[Allocate memory for record with environment capability, code capability, and acitvation code] Allocate a new piece of memory with room for a capability for the environment
+  \item[Store register contents to environment] Store the contents of each of the registers $r_1,\dots,r_n$ to the newly allocated memory.
+  \item[Allocate memory for record with environment capability, code capability, and acitvation code] Allocate a new piece of memory with room for a capability for the environment.
   \item[Store capabilities and activation code to record] Store the environment capability and code capability in the record followed by the activation code. 
-  \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the acitvation code and restrict it to an enter capability.
+  \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the acitvation code and restrict it to a global enter-capability.
   \item[Activation code:] \forcenewline
@@
 \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
-\item $\ms_\flag = [flag \mapsto 0]$
+\item $\ms_\flag = [\flag \mapsto 0]$
 \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
@@
     c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
-    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
     \reg & \in \Regs \\
@@
   \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
-  \item $\ms_\flag = [flag \mapsto 0]$
+  \item $\ms_\flag = [\flag \mapsto 0]$
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
@@
     c_{f3} & \defeq ((\rwx,\glob),\mathtt{f3}-\olf,\mathtt{3f},\mathtt{f3}) \\
-    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_{\adv},\start_\stk-1) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
     \reg & \in \Regs \\
@@
   \item $\ms_{f3}(\mathtt{f3}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f3}(\mathtt{f3}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f3}$ contains the code of $f3$.
-  \item $\ms_\flag = [flag \mapsto 0]$
+  \item $\ms_\flag = [\flag \mapsto 0]$
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
@@
 \end{verbatim}
+\begin{lstlisting}
+g1:  malloc $r_2$ 1
+     store $r_2$ 0
+     move $\pcreg$ $r_3$
+     lea $r_3$  $\dots$
+     crtcls $[(x, r_2)]$ $r_3$
+     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
+1g:  jmp $r_0$
+f4:  reqglob $r1$
+     reqperm $r_\stk$ $\encodePerm(\rwlx)$
+     store $x$ 0
+     scall $r_1$($[]$,$[r_0,r_1]$)
+     store $x$ 1
+     scall $r_1$($[]$,$[r_0]$)
+     load $r_1$ x
+     assert $r_1$ 1
+     cclear $\RegName \setminus \{r_0,\pcreg\}$
+4f:  jmp $r_0$
+\end{lstlisting}
+Where the $\dots$ is the appropriate offset to make the capability point to \texttt{f4}.
+
+\begin{lemma}[Correctness of $g1$]
+      \label{lem:correctness-g1}
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{g1} & \defeq ((\entry,\local),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
+    c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    \reg & \in \Regs \\
+    m & \defeq \hs_{g1} \uplus 
+        \ms_\flag \uplus                
+        \ms_\link \uplus 
+        \ms_\adv \uplus 
+        \ms_\malloc \uplus 
+        \ms_\stk \uplus
+        \ms_{\var{frame}} 
+  \end{align*}
+  where 
+  \begin{align*}
+    &\dom(\hs_{g1}) = [\mathtt{g1}-\olf,\mathtt{4f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link]\\
+    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $\ms_{g1}(\mathtt{g1}-\olf) = ((\readonly,\glob),\link,\link,\link)$, $\ms_{g1}(\mathtt{g1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{g1}$ contains the code of $g1$ immediately followed by the code of $f4$.
+  \item $\ms_\flag = [\flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
+  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of instruction.
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$}
 

### 2016-12-07 local-capability-machine.tex
Add a missing title.
g1 proof in progress. Made changes to f4. Added lemma bout nsim pwl and nequality.

@@
 \newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
+\renewcommand{\nsim}[1][n]{\stackrel{\tiny{#1}}{\simeq}}
 
@@
 \newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
-\newcommand{\entryCond}[1]{\plainfun{entryCondition}{#1}}
+\newcommand{\entryCond}[1]{\plainfun{enterCondition}{#1}}
 
@@
 
-      For convenience define let $i_1 \neq i_2$ and $i_1,i_2 \not\in \dom(\revokeTemp(W'))$ and define
+      For convenience define let $i_1 \neq i_2$ and $i_1,i_2 \not\in \dom(\revokeTemp{W'})$ and define
       \[
@@
       \[
-        \npair[n'']{\start_\adv,\addrend_\adv} \in \readCond(\glob)(W^{(3)})
+        \npair[n'']{\start_\adv,\addrend_\adv} \in \readCond{}(\glob)(W^{(3)})
       \]
@@
      store $x$ 0
-     scall $r_1$($[]$,$[r_0,r_1]$)
+     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
      store $x$ 1
-     scall $r_1$($[]$,$[r_0]$)
+     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
      load $r_1$ x
      assert $r_1$ 1
+     mclear $r_\stk$
      cclear $\RegName \setminus \{r_0,\pcreg\}$
@@
 \end{lstlisting}
+\lau{todo: when the stack is received the first time, make sure it points to where we would expect it to point (1 - range of auth.) Make macro for this.}
 Where the $\dots$ is the appropriate offset to make the capability point to \texttt{f4}.
@@
     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
-    c_{g1} & \defeq ((\entry,\local),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
+    c_{g1} & \defeq ((\entry,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
     c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
     c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
-    \reg & \in \Regs \\
     m & \defeq \hs_{g1} \uplus 
@@
   \[
-    (\reg\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
+    (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
   \]
@@
 \end{lemma}
-\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$}
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1}]
+  Let $n$ be given and make the remaining assumptions of the lemma. Define
+  \begin{align*}
+    W = & [0 \mapsto \iota^\sta (\perma,\ms_{g1})]\\
+        & [1 \mapsto \iota^\sta (\perma,\ms_\flag)]\\
+        & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
+        & [3 \mapsto \iota^\pwl]\\
+        & [4 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}] \\
+        & [5 \mapsto \iota_{\malloc,0}]
+  \end{align*}
+  \[
+    \ms = \ms_{g1} \uplus 
+        \ms_\flag \uplus                
+        \ms_\link \uplus 
+        \ms_\adv \uplus 
+        \ms_\malloc \uplus 
+        \ms_\stk
+  \]
+  and use the FTLR on $c_\adv$ (possible due to lemma~\ref{lem:nwlp-subset-pwl} and the chose world). From this, we get
+  \[
+    \npair[n+1]{c_\adv} \in \stder(W).
+  \]
+  Assuming we can show $\npair[n+1]{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$ and $\memSat[n+1]{\ms}{W}$, then we know
+  \[
+    \npair[n+1]{(\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}\update{\pcreg}{c_\adv},\ms) \in \observations(W)}
+  \]
+  Using $\ms_{\var{frame}}$ and resulting memory $m$ and the assumption from the lemma that
+  \[
+    (\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}\update{\pcreg}{c_\adv},m) \step[n] (\halted,m')
+  \]
+  We get some $W_{\var{final}} \futurestr W$, $\ms'$, and $\ms_r$ such that
+  \begin{align*}
+    & m' = \ms' \uplus \ms_r \uplus \ms_{\var{frame}}\\
+    & \memSat[1]{\ms'}{W_{\var{final}}}
+  \end{align*}
+  From island 1 (the flag island), the above memory segment satisfaction, and the fact that $W_{\var{final}} \futurestr W$ we get that part of $\ms'$ must be $\ms_\flag$ and thus
+  \[
+    m'(\flag) = 0
+  \]
+  which is the desired result. Now it suffices to show:
+  \begin{enumproof}
+    \item $\memSat[n+1]{\ms}{W}$
+      %TODO - this should be similar to previous proofs
+    \item $\npair[n+1]{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$
+      \begin{enumproof}
+        \item Case $r \in \RegName \setminus \{\pcreg,r_\stk,r_1,\}$: amounts to $\npair[n+1]{0} \in \stdvr(W)$, trivial.
+        \item Case $\npair[n+1]{c_\stk} \in \stdvr(W)$:
+          % TODO - same as in previous proof.
+        \item Case $\npair[n+1]{c_{g1}} \in \stdvr(W)$:\\
+          let $n' < n+1$ and $W' \futurestr W$ be given and show
+          \[
+            \npair[n']{\updatePcPerm{c_{g1}}} \in \stder(W')
+          \]
+          to this end let $n'' \leq n'$, $\ms'$, and $\reg'$ be given and assume
+          \begin{align}
+            & \npair[n'']{\reg'} \in \stdrr(W') \label{g1:regrel1} \\
+            & \memSat[n'']{\ms'}{W'} \label{g1:memsat1}
+          \end{align}
+          and show
+          \[
+            \npair[n'']{(\reg'\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms')} \in \observations(W')
+          \]
+          To this end let $i \leq n''$, $\ms_f$, $m''$ be given and assume
+          \begin{equation}
+            (\reg'\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms' \uplus \ms_f) \step[i] (\halted,m'') \label{g1:execution}
+          \end{equation}
+          and show
+          \begin{equation}
+            \exists W_{\var{str}}' \futurestr W', ms_r, \ms'' \ldotp m'' = \ms'' \uplus \ms_r \uplus \ms_f \land \memSat[n''-i]{\ms''}{W_{\var{str}}'} \label{g1:g-goal}
+          \end{equation}
+          consider how the above execution goes. By \ref{g1:memsat1} we know that the instructions of $g1$ are unaltered. The following happens
+          \begin{itemize}
+            \item ...
+            %TODO - typeout the details. 
+          \end{itemize}
+          In the end, we have a new memory segment
+          \[
+            \ms'' = \ms_\malloc' \uplus \ms_x \uplus \ms_{\var{env}} \uplus \ms_{\var{cls}}
+          \]
+          where $\ms_x(x) = 0$, $\ms_{\var{env}}(\var{env}) = ((\rwx,\glob),x,x,x)$, and $\ms_{\var{cls}}=[\var{cls} \mapsto c_{\var{env}}, \var{cls} + 1 \mapsto c_{f4}, \var{cls} + 2 \mapsto i^{ac}_1, \var{cls} + m + 1 \mapsto i^{ac}_m]$ where $i^{ac}_1, \dots, i^{ac}_m$ are the instructions of the activation code, $c_{\var{env}} = ((\rwx,\glob),\var{env},\var{env},\var{env})$, $c_{f4} = ((\exec,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{f4})$.
+          Further at this point, the register file contains:
+          \[
+            \reg'' = reg_0\update{\pcreg}{\updatePcPerm{\reg'(r_0)}}\update{r_0}{\reg'(r_0)}\update{r_1}{((\entry,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)}
+          \]
+          We know that $\npair[n'']{\reg'(r_0)} \in \stdvr(W')$.
+We have assumed execution does not fail, so $\reg'(r_0)$ must be a capability with permission $\entry$, $\exec$, $\rwx$, or $\rwlx$.
+In either case, we get $\execCond{}$ or $\entryCond{}$ (for certain parameters).
+Now use $n^{(3)}$ ($n''$ minus the steps the two malloc calls used) and $W'' \futurewk W'$ where
+          \[
+            W'' = W'\update{5}{\iota_\malloc'}
+                    \update{i_1}{\iota^\sta (\perm,\ms_{\var{env}})}
+                    \update{i_2}{\iota^\sta (\perm,\ms_{\var{cls}})}
+                    \update{i_3}{\iota_x}
+          \]
+          where $i_1,i_2,i_3 \not\in\dom(W')$ and they distinct, and $\iota_\malloc \futurewk W(5)$ and
+          \begin{align*}
+            \iota_x & = (\perm,0,\{(0,1),(0,0),(1,1)\},\{(1,0),(0,1),(0,0),(1,1)\},H_x) \\
+            H_x \; s \; W & = \{\npair{\ms} \mid \ms(x) = s \land \dom(\ms) = [x,x]\}
+          \end{align*}
+          %TODO insert tikz picture of machine (just like Dreyer Neis machine.)
+          with $\execCond{}$ or $\entryCond{}$ to get
+          \[
+            \npair[n^{(3)}]{\updatePcPerm{\reg'(r_0)}} \in \stder(W'')
+          \]
+          If we can show $\npair[n^{(3)}]{\reg''} \in \stdrr(W'')$ and $\memSat[n^{(3)}]{\ms''}$, then using assumption \ref{g1:execution} we can get \ref{g1:g-goal}.
+
+          SFTS:
+          \begin{enumproof}
+            \item$\memSat[n^{(3)}]{\ms''}$
+              %TODO should be simple - all existing regions have interpretations monotone wrt public future worlds.
+The malloc island uses its "isolation condition" which makes room for the two new islands.
+The x islands starts in state 0 (and the appropriate memory address contains 0).
+The env and cls islands are static islands designed to accept exactly what the memory contains.
+            \item $\npair[n^{(3)}]{\reg''} \in \stdrr(W'')$ amounts to showing:
+              \begin{enumproof}
+                \item For $r \in \RegName \setminus \{\pcreg, r_0, r_1 \}$ $\reg''(r) = 0$ and it is trivial to show $\npair[n^{(3)}]{0} \in \stdvr(W'')$. 
+                \item $\npair[n^{(3)}]{\reg'(r_0)} \in \stdvr(W'')$ - follows from \ref{g1:regrel1} and public monotonicity of $\stdvr$.
+                \item $\npair[n^{(3)}]{((\entry,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)} \in \stdvr(W'')$ \\
+                  To this end let $n^{(4)} < n^{(3)}$, $W^{(3)} \futurestr W''$ be given and assume
+                  \begin{align*}
+                    \\ \npair[n^{(4)}]{\reg^{(3)}} \in \stdrr(W^{(3)}) &
+                    \\ \memSat[n^{(4)}]{\ms^{(3)}}{W^{(3)}} \label{g1:memsat-cls}
+                  \end{align*}
+                  for some $\reg^{(3)}$ and $\ms^{(3)}$ and show
+                  \[
+                    \npair[n^{(4)}]{(\reg^{(3)}\update{\pcreg}{((\exec,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)},\ms^{(3)})} \in \observations(W^{(3)})
+                  \]
+                  to this end let $j \leq n^{(4)}$, $m^{(3)}$, and $m_f'$ be given and assume
+                  \begin{equation}
+                    (\reg^{(3)}\update{\pcreg}{((\exec,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)},\ms^{(3)} \uplus \ms_f') \step[j] (\halted, m^{(3)}) \label{g1:execution2}
+                  \end{equation}
+                  and show
+                  \begin{equation}
+                    \exists W_{\var{str}}^{(3)} \futurestr W^{(3)}, \ms'^{(3)}, \ms_r \ldotp m^{(3)} = \ms'^{(3)} \uplus \ms_r \uplus \ms_f' \land \memSat[n^{(4)}-j]{\ms'^{(3)}}{W_{\var{str}}'} \label{g1:g-goal2}
+                  \end{equation}
+                  to this end consider what happens in execution \ref{g1:execution2}: Due to \label{g1:memsat-cls} we know that the activation code and $f4$ code is unaltered.
+Running the activation code loads the environment capability to $r_{\var{env}}$ and loads the $f4$ capability, then the $r_1$ register is tested whether it contains a global capability and the $r_\stk$ register is tested for whether it contains an $\rwlx$ capability, say $\reg''(r_\stk) = ((\rwlx, \gl), \start_{\stk}', \addrend_{\stk}', \addr_{\stk}')$.
+As we know the execution halts succesfully, both of these checks must succeed.
+
+                  Continuing execution until just after the point where we jump to the capability in register $r_1$ will store $0$ to address $x$ of memory, push restore code, $\reg^{(3)}(r_0)$, $\reg^{(3)}(r_1)$, push $\reg^{(3)}(r_\stk)$, and a capability to return to in $f4$.
+
+                  The memory and register file at this point are as follows: %TODO
+                  \[
+                  \dots  
+                  \]
+
+                  We have assumed that the execution does not fail which entails several things: The capability in $r_1$ is executable after being jumped to.
+The capability $r_\stk$ has authority over some none empty part of the memory.
+
+                  The latter is used with Lemma~\ref{lem:pwl-nsim-view}
+
+                  As $\reg^{(3)}(r_\stk) = ((\rwlx,\gl),\start,\addrend,\addr)$ for some $\gl$, $\start$, $\addrend$, and $\addr$ and $\npair[n^{(4)}]{\reg^{(3)}(r_\stk)} \in \stdvr(W^{(3)})$ we know that
+                  \begin{align*}
+                    \\ \npair[n^{(4)}]{(\start,\addrend)} \in \readCond{}(\gl)(W) &
+                    \\ \npair[n^{(4)}]{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
+                  \end{align*}
+                  which means that for some region $r$ and interval $[\start',\addrend'] \supseteq [\start,\addrend]$ we have 
+                  \[
+                    W(r) \nsim[n^{(4)}] \iota^\pwl_{\start',\addrend'}
+                  \]
+
+              \end{enumproof}
+          \end{enumproof}
+      \end{enumproof}
+  \end{enumproof}
+\end{proof}
+
+\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
 
@@
     &\forall n,W_1,W_2,W_1' \ldotp
-    &\quad W_1 \nequal W_2 \land W_1' \futurestr W_1 \Rightarrow \exists W_2' \ldotp W_2' \nequal W_1' \land W_2' \futurestr W_2
+    &\quad W_1 \nequal W_2 \land W_1' \futurestr W_1 \Rightarrow \exists W_2' \ldotp W_2' \nequal W_1' \land W' \futurestr W_2
   \end{align*}
@@
 \end{lemma}
+
+\begin{lemma}
+  \label{lem:pwl-nsim-view}
+  \begin{align*}
+    & \forall n \in \nats \ldotp \forall \start, \addrend, b \in \Addrs \ldotp \forall \iota \in \Regions \\
+    & \iota \nsim \iota^\pwl_{\start,\addrend} \land a \leq b \Rightarrow \iota \nequal \iota^\pwl_{\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:pwl-nsim-view}]
+  For $n = 0$ it is trivial, so assume $n > 0$.
+  Say $\iota = (v,s,\phi_{\var{pub}},\phi,H)$, then by $\nsim$, we know $s = 1$, $\phi_{\var{pub}} \equiv \phi \equiv =$, and $H = H^\pwl_{\start,\addrend}$.
+It remains to show that $v=\temp$.
+To do so, we show that it cannot be the case that $v=\perma$.
+If $v = \perma$, then $H$ must be monotone with respect to $\futurestr$.
+If we can show that this is not the case, then for $\iota$ to be a region it must be the case that $v \neq \perma$ and thus $v = \temp$.
+
+To this end let $b\not\in[\start,\addrend]$ and define the worlds:
+\begin{align*}
+  \xi (W)  =& [0 \mapsto \iota^\pwl_{\start,\addrend}]  \\ 
+      & [1 \mapsto \iota^\pwl_{b,b}] \\
+  \xi (W')  =& [0 \mapsto \iota^\pwl_{\start,\addrend}]  \\ 
+      & [1 \mapsto \revoked] 
+\end{align*}
+For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$ $W' \futurestr W$. Now define the following memory segment:
+\[
+  \ms = [\start \mapsto ((\readonly,\local),b,b,b), \start+1 \mapsto 0, \dots , \addrend \mapsto 0]
+\]
+It is the case that
+\[
+  \npair{\ms} \in H \; 1 \; W
+\]
+but
+\[
+  \npair{\ms} \not\in H \; 1 \; W'
+\]
+as it is not the case that \lau{what if $n=1$? Need to figure out when exactly this holds}
+\[
+  \npair[n-1]{((\readonly,\local),b,b,b)} \in \stdvr(\xi(W')).
+\]
+The only other option that remains is $v=\temp$.
+\end{proof}
+\lau{$a \leq b$ that the interpretation is not empty.}
 

### 2016-12-07 local-capability-machine.tex
Added prepstack macro and used it in f4 program

@@
 \item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with $\decodePerm{n}$. If not fail, otherwise continue execution.
+\item[\texttt{prepstack} $r$] Tests if register $r$ contains a capability with $\decodePerm{n}$.
+If not fail, otherwise assume $r$ points to $((\rwlx,\gl),\start,\addrend,\addr)$ adjust it to $((\rwlx,\gl),\start,\addrend,\start - 1)$.
+\lau{What if the stack starts at address 0? Changing the stack convention to always point at the first free cell will get rid of this problem (because our memory us uncapped.)}
 \item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 
@@
 \item The code of a closure will often be found in conjunction with the code that creates it.
+\item \texttt{prepstack} as ``prepare stacl''. This ensures that the register contains something that looks like a stack and it is prepared for our stack convention.
 \end{itemize}
@@
 f4:  reqglob $r1$
-     reqperm $r_\stk$ $\encodePerm(\rwlx)$
+     prepstack $r_\stk$
      store $x$ 0
@@
 \end{lstlisting}
-\lau{todo: when the stack is received the first time, make sure it points to where we would expect it to point (1 - range of auth.) Make macro for this.}
 Where the $\dots$ is the appropriate offset to make the capability point to \texttt{f4}.
@@
 \end{mathpar}
+
+\subsection{The logical relation}
 

### 2016-12-08 local-capability-machine.tex
push with pwl lemma, awkward example work in progress

@@
 
-%Math notation
+% Math notation
 \newcommand{\restrictfun}[1]{|_{#1}}
@@
 
-%cofes
+% cofes
 \newcommand{\cofe}{c.o.f.e.}
@@
 
-%Comments
+% Comments
 \newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
@@
 
-%Variables
+% Variables
 \newcommand{\var}[1]{\mathit{#1}}
@@
 
-%Memory projections
+% Memory projections
 \newcommand{\plainproj}[1]{\mathrm{#1}}
@@
 
-%Configuration end states
+% Configuration end states
 \newcommand{\failed}{\textsl{failed}}
@@
 
-%Functions
+% Functions
 \newcommand{\plainfun}[2]{
@@
 
-%World operations
+% World operations
 \newcommand{\future}{\mathbin{\sqsupseteq}}
@@
 
-%Assembly labels
+% Assembly labels
 \newcommand{\codelabel}[1]{\mathit{#1}}
@@
 
-%Type(s)
+% Type(s)
 \newcommand{\type}[1]{\mathrm{#1}}
@@
 
-%Domains
+% Domains
 \newcommand{\plaindom}[1]{\mathrm{#1}}
@@
 
-%LR
+% LR
 \newcommand{\intr}[2]{\mathcal{#1}}
@@
 
-%Reference register/memory
+% Reference register/memory
 \newcommand{\refreg}[1]{\lfloor #1 \rfloor}
@@
 
-%Instructions
-%No arguments
+% Instructions
+% No arguments
 \newcommand{\zinstr}[1]{\mathtt{#1}}
@@
 \newcommand{\halt}{\zinstr{halt}}
-%One argument
+% One argument
 \newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
 \newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
-%Two arguments
+% Two arguments
 \newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
@@
 \newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
-%Three arguments
+% Three arguments
 \newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
@@
 
-%Permissions
+% Permissions
 \newcommand{\plainperm}[1]{\mathrm{#1}}
@@
 \newcommand{\rwx}{\plainperm{rwx}}
-%PWL permissions
+% PWL permissions
 \newcommand{\readwritel}{\plainperm{rwl}}
@@
 
-%Global/local
+% Global/local
 \newcommand{\local}{\plainperm{local}}
@@
 
-%Views
+% Views
 \newcommand{\plainview}[1]{\mathrm{#1}}
@@
 
-%OP sem
+% OP sem
 \newcommand{\diverge}[1][n]{\not\Downarrow_{#1}}
@@
 \end{align*}
-%TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
+% TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
 \begin{align*}
@@
                 \end{tabular}
-\caption{ Stack and register-register file after the clean up code has been run. The ``private'' words have been popped to their respective registers. The restore code has been popped off the stack. }
+                \caption{ Stack and register-register file after the clean up code has been run.
+The ``private'' words have been popped to their respective registers.
+The restore code has been popped off the stack.
+}
               \end{figure}
@@
               \subsubsection{Simple use of local capabilities}
-The following programs only use local capabilities when they create protected return pointers. The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
+              The following programs only use local capabilities when they create protected return pointers.
+The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
 
@@
                 \item $\memSat[n+1]{\ms}{W}$
-      %TODO - this should be similar to previous proofs
+                  % TODO - this should be similar to previous proofs
                 \item $\npair[n+1]{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$
@@
                     \item ...
-            %TODO - typeout the details. 
+                      % TODO - typeout the details. 
                     \end{itemize}
@@
                       \update{i_2}{\iota^\sta (\perm,\ms_{\var{cls}})}
-                    \update{i_3}{\iota_x}
+                      \update{i_3}{\iota_{x,0}}
                     \]
@@
                     \begin{align*}
-            \iota_x & = (\perm,0,\{(0,1),(0,0),(1,1)\},\{(1,0),(0,1),(0,0),(1,1)\},H_x) \\
+                      \iota_{x,s} & = (\perm,s,\{(0,1),(0,0),(1,1)\},\{(1,0),(0,1),(0,0),(1,1)\},H_x) \\
                       H_x \; s \; W & = \{\npair{\ms} \mid \ms(x) = s \land \dom(\ms) = [x,x]\}
                     \end{align*}
-          %TODO insert tikz picture of machine (just like Dreyer Neis machine.)
+                    % TODO insert tikz picture of machine (just like Dreyer Neis machine.) and prove that it is indeed a region (should be simple, interpretation does not depend on world).
                     with $\execCond{}$ or $\entryCond{}$ to get
@@
                     \item$\memSat[n^{(3)}]{\ms''}$
-              %TODO should be simple - all existing regions have interpretations monotone wrt public future worlds.
+                      % TODO should be simple - all existing regions have interpretations monotone wrt public future worlds.
 The malloc island uses its "isolation condition" which makes room for the two new islands.
@@
                         To this end let $n^{(4)} < n^{(3)}$, $W^{(3)} \futurestr W''$ be given and assume
-                  \begin{align*}
-                    \\ \npair[n^{(4)}]{\reg^{(3)}} \in \stdrr(W^{(3)}) &
+                        \begin{align}
+                          \\ \npair[n^{(4)}]{\reg^{(3)}} \in \stdrr(W^{(3)}) \label{g1:reg-file-cls} &
                           \\ \memSat[n^{(4)}]{\ms^{(3)}}{W^{(3)}} \label{g1:memsat-cls}
-                  \end{align*}
+                        \end{align}
                         for some $\reg^{(3)}$ and $\ms^{(3)}$ and show
@@
                         to this end consider what happens in execution \ref{g1:execution2}: Due to \label{g1:memsat-cls} we know that the activation code and $f4$ code is unaltered.
-Running the activation code loads the environment capability to $r_{\var{env}}$ and loads the $f4$ capability, then the $r_1$ register is tested whether it contains a global capability and the $r_\stk$ register is tested for whether it contains an $\rwlx$ capability, say $\reg''(r_\stk) = ((\rwlx, \gl), \start_{\stk}', \addrend_{\stk}', \addr_{\stk}')$.
-As we know the execution halts succesfully, both of these checks must succeed.
+Running the activation code loads the environment capability to $r_{\var{env}}$ and loads the $f4$ capability, then the $r_1$ register is tested whether it contains a global capability (which it does s execution does not fail) and the $r_\stk$ register is tested for whether it contains an $\rwlx$ capability (which it does as execution does not fail), and the pointer address is adjusted, so the $r_\stk$-register now contains $((\rwlx, \gl), \start_{\stk}', \addrend_{\stk}', \start_{\stk}'-1)$.
 
                         Continuing execution until just after the point where we jump to the capability in register $r_1$ will store $0$ to address $x$ of memory, push restore code, $\reg^{(3)}(r_0)$, $\reg^{(3)}(r_1)$, push $\reg^{(3)}(r_\stk)$, and a capability to return to in $f4$.
+
+                        Now use \ref{g1:memsat-cls} to get $P$ such that $\memSat[n^{(4)}]{\ms^{(3)}}{W^{(3)}}{P}$. Now use Lemma~\ref{lem:priv-mono-like2} to get $\ms^{(4)}$ and $\ms_R$ such that
+                        \begin{align*}
+                          & \ms^{(3)} = \ms^{(4)} \uplus \ms_R \\
+                          & \memSat[n^{(4)}]{\ms^{(4)}}{\revokeTemp{W^{(3)}}}{P|_{\dom(\erase{W^{(3)}}{\perma})}} \\
+                          & \ms_R = \biguplus_{r \in \erase{W^{(3)}}{\temp}} P(r) \land \\
+                          & \ms^{(4)} = \biguplus_{r \in \erase{W^{(3)}}{\perma}} P(r)
+                        \end{align*}
+                        Now we want to argue that the ``stack'' we received is part of the revoked part.
+To do so, it suffices to show that it is governed by a temporary region.
+From \ref{g1:reg-file-cls} and $\reg^{(3)}(r_\stk)$ is a $\rwlx$ capability, we get
+                        \begin{align*}
+                          &\npair[n^{(4)}]{\start_\stk',\addrend_\stk'} \in \readCond{}(\gl)(W^{(3)}) \\
+                          &\npair[n^{(4)}]{\start_\stk',\addrend_\stk'} \in \writeCond{}(\iota^\pwl,\gl)(W^{(3)})
+                        \end{align*}
+                        from which we can infer that for some $\start_\stk''$, $\addrend_\stk'$, and $r\in\dom(W^{(3)})$
+                        \[
+                          W^{(3)}(r) \nsim[n^{(4)}-1] \iota^\pwl_{\start_\stk'',\addrend_\stk''}
+                        \]
+                        from the fact that execution does not fail, we can infer $\start_\stk'' \leq \addrend_\stk''$. Now use Lemma~\ref{lem:pwl-nsim-view} to get
+                        \[
+                          W^{(3)}(r) \nequal[n^{(4)}-1] \iota^\pwl_{\start_\stk'',\addrend_\stk''}
+                        \]
+                        from which we can infer that the view of $W^{(3)}(r)$ is $\temp$.
+
+                        We can from this infer that $\ms^{(3)} |_{\start_\stk'',\addrend_\stk''} \subset \ms_R$.
+
+                        Consider the changes made to memory: $x$ was set to 0, so we have $\ms^{(4)}[x \mapsto 0]$.$W^{(3)}(i_3) \futurestr W''(i_3)$, so $W^{(3)}(i_3) \in \{0,1\}$.
+Now define $W^{(3)}_0 = \revokeTemp{W^{(3)}}[i_3 \mapsto \iota_{x,0}]$ we have $W^{(3)}_0 \futurestr W^{(3)}$.
+Further
+                        \[
+                          \memSat[n^{(4)}]{\ms^{(4)}[x \mapsto 0]}{W^{(3)}_0}{P|_{\dom(\erase{W^{(3)}}{\perma})}[i_3 \mapsto [x \mapsto 0]]}
+                        \]
+                        which follows from $H$ of $\perma$ regions being mono wrt $\futurestr$ and $x$ mapping to 0.
+
+                        Define 
+                        \[
+                          W_{\var{reused}} = [i_1' \mapsto \iota^\sta (\temp,\ms|_{\start_\stk'',\start_\stk'''-1})][i_2' \mapsto \iota^\pwl_{\start_\stk''',\addrend_\stk''}]
+                        \]
+                        for $i_1',i_2' \not\in \dom(W^{(3)})$ and distinct where $\start_\stk'''$ is the first address of the stack we are passing on.
+
+                        It is the case that
+
+                        % Stack and x are the only things to change in memory.
+
+                        % TODO: continue here.
 
@@
 
-\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
+              \lau{If $c_{g1}$ is global it seems like we will run into trouble.
+When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
 
@@
                 \begin{align*}
-    & \forall \ms, n, W, R : \activeReg(W) \fun \MemSegments  \ldotp \\
+                  & \forall \ms, n, W, R : \activeReg{W} \fun \MemSegments  \ldotp \\
                   & \quad \memSatPar{\ms}{W}{P} \Rightarrow \\
@@
 
+              \begin{lemma}[Revoke temporary memory with stack]
+                \label{lem:revoke-temp-stack}
+                \begin{align*}
+                  & \forall n, \ms, W, \reg, r_\stk,\gl,\start,\addrend,\addr \ldotp\\
+                  & \quad \memSat{\ms}{W} \land \npair{\reg} \in \stdrr(W) \land \\
+                  & \quad \reg(r_\stk) = ((\rwlx,\gl),\start,\addrend,\addr) \land b \leq e \\
+                  & \qquad \exists \ms', \ms_r \ldotp \\
+                  & \qquad \quad \memSat{\ms }{\revokeTemp{W}} \land \ms = \ms' \uplus \ms_r
+                \end{align*}
+              \end{lemma}
+
               \begin{lemma}[Disjoint memory satisfaction]
@@
                   & \forall n \in \nats \ldotp \forall \start, \addrend, b \in \Addrs \ldotp \forall \iota \in \Regions \\
-    & \iota \nsim \iota^\pwl_{\start,\addrend} \land a \leq b \Rightarrow \iota \nequal \iota^\pwl_{\start,\addrend}
+                  & \iota \nsim \iota^\pwl_{\start,\addrend} \land \start \leq \addrend \Rightarrow \iota \nequal \iota^\pwl_{\start,\addrend}
                 \end{align*}
@@
                 \end{align*}
-For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$ $W' \futurestr W$. Now define the following memory segment:
+                For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$, we have $W' \futurestr W$. Now define the following memory segment:
                 \[

### 2016-12-09 local-capability-machine.tex
indentation g1
indentation

@@
                       \item Show $\forall a \in [\start_\adv,\addrend_\adv] \ldotp \npair[n'-1]{\ms(a)} \in \stdvr(\revokeTemp{W})$
-                      \item Notice: $W = \revokeTemp(W)$
+                      \item Notice: $W = \revokeTemp{W}$
                       \end{enumproof}
@@
       \[
-                      \npair[n']{\updatePcPerm(c_\prp)} \in \stder(W')
+        \npair[n']{\updatePcPerm{c_\prp}} \in \stder(W')
       \]
@@
       \[
-                      \npair[n'']{(\reg\update{\pcreg}{\updatePcPerm(c_\prp)},\ms)}
+        \npair[n'']{(\reg\update{\pcreg}{\updatePcPerm{c_\prp}},\ms)} \in \observations(W')
       \]
@@
 
-              \lau{If $c_{g1}$ is global it seems like we will run into trouble.
-When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
+\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
 
@@
 
+\begin{lemma}
+  Say the instructions for \texttt{scall} $r([r_{j_1},\dots,r_{j_n}],[r_{k_1},\dots, r_{k_m}])$ are $i_0,\dots,i_l$, then
+  \begin{align*}
+    & m = \ms \uplus \ms_\stk \uplus \ms_f \\
+    & \memSat{\ms \uplus \ms_\stk}{W}
+    & \npair{\reg[\pc \mapsto c_{\var{scall}}][r_\stk \mapsto c_\stk]} \in \stdrr(W) \\
+    & \ms_\stk = m | _{\start_\stk,\addrend_\stk} \\
+    & c_\stk = ((\rwlx,\gl),\start_\stk,\addrend_\stk,\addr) \\
+    & \start_\stk-1 \leq \addr \land \start_\stk \leq \addrend_\stk \land \addr \leq \addrend_\stk \\
+    & c_{\var{scall}} = ((\perm,\gl)\start,\addrend,\addr) \\
+    & \perm \in \{\exec, \rwx, \rwlx\} \\
+    & \addrend - \addr \leq l  \\
+    & \ms(\addr) = i_0, \dots, \ms(\addr+l) = \mathtt{jmp r}
+    & (\reg[\pc \mapsto c_{\var{scall}}][r_\stk \mapsto c_\stk],m) \step[i'] (\reg_0[r_{k_1} \mapsto \reg(r_{k_1})]\dots[r_{k_m} \mapsto \reg(r_{k_m})][\pc \mapsto c_{\var{scall}}'][r_\stk \mapsto c_\stk'],m) \\
+    & c_{\var{scall}}'.\mathrm{a} = \addr+l \\
+    & \quad \Rightarrow \\
+    & \qquad \exists \ms_\stk', \ms', \ms_r \ldotp \\
+    & \qquad \quad W' = \revokeTemp{W}[\var{rn}_1\mapsto\iota^\sta(\temp, )][\var{rn}_2 \mapsto \iota^\pwl] \\
+    & \qquad \quad \memSat[n-i']{\ms_\stk'\uplus \ms'}{W'}
+  \end{align*}
+\end{lemma}
+
 \section{Other examples and applications}

### 2016-12-09 local-capability-machine.tex
more indentation
.
Tentative formulation of an scall lemma.
changed conclusion of scall lemma to be in terms of observation relation
scall lemma: changed confs to conform with domain
removed preview from preamble

@@
 %% make latex preview-mode work with natbib...
-\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
+%\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
 
@@
 \subsection{Lemmas}
+
+\begin{definition}
+  We say that $(\ms,\reg) \text{ is looking at } [i_0,\cdots,i_n]$ if
+  \begin{itemize}
+  \item $\reg(\pcreg) = ((p,g),b,e,a)$
+  \item $p = \rwx$, $p = \exec$, or $p = \rwlx$
+  \item $a+n\leq e$, $b\leq a\leq e$
+  \item $\ms(a+0,\cdots,a+n) = [i_0,\cdots,i_n]$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}
+  $(\ms,\reg) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
+
+  \begin{itemize}
+  \item $\reg(r_\stk) =((\rwlx,\_),b_\stk,e_\stk,a_\stk)$
+  \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
+  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$
+  \item $b_\stk - 1\leq a_\stk$
+  \end{itemize}
+\end{definition}
+
+\begin{lemma}[$\mathtt{scall}$ works]
+  If
+  \begin{itemize}
+  \item $\heapSat[\ms]{n}{W}$ 
+  \item $(\reg,\ms) \text{ is looking at } \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
+    \overline{r_{\mathit{priv}}}))$
+  \item $(\reg,\ms) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
+  \item[Hyp-Callee] If
+    \begin{itemize}
+    \item $\ms_{\mathit{unused}} = \ms_{\mathit{act}}
+      \uplus \ms_{\mathit{unused}}'$
+    \item $W' =
+      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
+    \item $\memSat[n-1]{\ms''}{W'}$
+    \item $(\reg',\ms'') \text{ has stack $\emptyset$ with
+        $\ms_{\mathit{unused}}'$ unused}$,
+    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
+      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
+      \mapsto c_{\mathit{ret}}]$
+    \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
+    \end{itemize}
+    then we have that $(\reg',\ms'') \in \observations(W')$
+  \item[Hyp-Cont] ...
+  \end{itemize}
+  Then 
+  \begin{itemize}
+    \item $(\reg,\ms) \in \observations(W)$
+  \end{itemize}
+
+  Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
+  words for stack pointer and return pointer and some fixed number of words for
+  the standard return instructions).
+\end{lemma}
+\begin{proof}
+  Apply Hyp-Callee with $\ms_{\mathit{act}} = ...$. 
+\end{proof}
+
 We expect the following lemmas to hold true:

### 2016-12-12 local-capability-machine.tex
added comment about old text
missing stk cap in scall lemma?

@@
 We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
+\lau{The above program and text are from before we introduced macros. They can be deleted when we have gotten used to the style of the new programs. }
 
@@
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
-  This proof goes like the proof of Lemma~\ref{lem:correctness-f3} except for item \ref{f2:c-prp}
+  This proof goes like the proof of Lemma~\ref{lem:correctness-f2} except for item \ref{f2:c-prp}
   \begin{enumproof}[start=3]
@@
       \[
-        \npair[n''] \in \stder(W^{(3)})
+        \npair[n'']{\updatePcPerm{c_\adv}} \in \stder(W^{(3)})
       \]
@@
       \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}]$
+      \mapsto c_{\mathit{ret}}]$ \lau{add stack capability?}
     \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$

### 2016-12-14 local-capability-machine.tex
typos
Further work on the tentative lemma about scall.
changed scall lemma and added fetch for linking

@@
 \newcommand{\readwritel}{\plainperm{rwl}}
+\newcommand{\rwl}{\readwritel}
 \newcommand{\rwlx}{\plainperm{rwlx}}
@@
               \begin{description}
-              \item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline
+              \item[\texttt{fetch} $r$ $f$] load the entry of the linking table corresponding to $f$ to register $r$.
+              \item[\texttt{call} $r(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline
                 $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
@@
                 \item [Clear unused registers]
-                  Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
-                \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
+                  Clear all the register that are not $\pcreg$, $r$, $r_0$ or in $\bar{r}_{\var{args}}$.
+                \item [Jump] Jump to $r$
                 \item [Activation code] The activation code does the following:
@@
 If they are unequal, then the assertion flag named $\var{flag}$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
-              \item[\texttt{cclear $r$}] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
+              \item[\texttt{mclear $r$}] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
 In a real system it would probably be implemented as a vector operation which allows modification of continuous segments of memory rather fast.}
@@
               \item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, and decrementing the address of the stack capability.
-              \item[\texttt{scall} \forcenewline
-                $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
+              \item[\texttt{scall} $r(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline$\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
 This call assumes $r_{\var{stk}}$ contains a stack capability.
@@
                 \item [Clear unused registers]
-                  Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
-                \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
+                  Clear all registers but $\pcreg$, $r$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.
+                \item [Jump] Jump to register $r$.
+\\ \lau{scall used to be dependent on linking, but now this part has been extracted to the fetch macro to make scall independent of linking.}
+%Load the entry $f$ from the linking table to a free register and jump to it.
                 \item [Run the on stack restore code]
@@
               \item[\texttt{reqglob $r$}] Tests if register $r$ contains a $\glob$ capability. If not fail, otherwise continue execution.
-              \item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with $\decodePerm{n}$. If not fail, otherwise continue execution.
-              \item[\texttt{prepstack} $r$] Tests if register $r$ contains a capability with $\decodePerm{n}$.
+              \item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with permission $\decodePerm{n}$. If not fail, otherwise continue execution.
+              \item[\texttt{prepstack} $r$] Tests if register $r$ contains a capability with permission $\rwlx$.
 If not fail, otherwise assume $r$ points to $((\rwlx,\gl),\start,\addrend,\addr)$ adjust it to $((\rwlx,\gl),\start,\addrend,\start - 1)$.
@@
               \item The code of a closure will often be found in conjunction with the code that creates it.
-              \item \texttt{prepstack} as ``prepare stacl''. This ensures that the register contains something that looks like a stack and it is prepared for our stack convention.
+              \item \texttt{prepstack} as ``prepare stack''. This ensures that the register contains something that looks like a stack and it is prepared for our stack convention.
               \end{itemize}
@@
                 \caption{This is the first figure of 6 that illustrates how \texttt{scall} works.
-In this example, the call \texttt{scall $f([r_{\var{args},1},\dots,r_{\var{args},n}],[r_0,r_{\var{priv},1},\dots,r_{\var{priv},m}])$.
+In this example, the call \texttt{scall $r([r_{\var{args},1},\dots,r_{\var{args},n}],[r_0,r_{\var{priv},1},\dots,r_{\var{priv},m}])$.
 In this example the two lists of registers are disjoint even though that does not have to be the case.}}
@@
 f2: push 1
-    scall adv([],[])
+    fetch r1 adv
+    scall r1([],[])
     pop r1
@@
 f3: push 1
-    scall adv([],[])
+    fetch r1 adv
+    scall r1([],[])
     pop r1
@@
     push 2
-    scall adv([],[])
+    fetch r1 adv
+    scall r1([],[])
 3f: halt
@@
      mclear $r_\stk$
-     cclear $\RegName \setminus \{r_0,\pcreg\}$
+     rclear $\RegName \setminus \{r_0,\pcreg\}$
 4f:  jmp $r_0$
@@
 \begin{definition}
-  We say that $(\ms,\reg) \text{ is looking at } [i_0,\cdots,i_n]$ if
+  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ if
   \begin{itemize}
@@
   \item $\ms(a+0,\cdots,a+n) = [i_0,\cdots,i_n]$
+  \item $c_{\mathit{next}} = ((p,g),b,e,a+n+1)$
   \end{itemize}
@@
 \begin{definition}
-  $(\ms,\reg) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
+  $(\reg,\ms) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
 
@@
   \item $\heapSat[\ms]{n}{W}$ 
-  \item $(\reg,\ms) \text{ is looking at } \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
-    \overline{r_{\mathit{priv}}}))$
+  \item $(\reg,\ms) \text{ is looking at }
+    \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
+    \overline{r_{\mathit{priv}}})) \text{ followed by } c_{\mathit{next}}$
   \item $(\reg,\ms) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
@@
     \item $\ms_{\mathit{unused}} = \ms_{\mathit{act}}
-      \uplus \ms_{\mathit{unused}}'$
+      \uplus \ms_{\mathit{unused}}'$,
     \item $W' =
@@
     \item $(\reg',\ms'') \text{ has stack $\emptyset$ with
-        $\ms_{\mathit{unused}}'$ unused}$,
+        $\ms_{\mathit{unused}}'$ unused}$
+    \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
     \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
       \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}]$ \lau{add stack capability?}
+      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_]$ 
     \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
     \end{itemize}
-    then we have that $(\reg',\ms'') \in \observations(W')$
-  \item[Hyp-Cont] ...
+    then we have that $(n-1,(\reg',\ms'')) \in \observations(W')$
+  \item[Hyp-Cont] If
+    \begin{itemize}
+    \item $n' \leq n-2$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)]$
+    \item for all $r$, we have that:
+      \begin{equation*}
+        \reg'(r)
+        \begin{cases}
+          = c_{\mathit{next}} &\text{ if } r = \pcreg\\
+          = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
+          \in \stdvr(W'') &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\}
+        \end{cases}
+      \end{equation*}
+    \item $(\reg',\ms'')$ has stack $\ms_\stk$ with
+        $\ms_{\mathit{unused}}''$ unused for some $\ms_{\mathit{unused}}''$
+    \item $\memSat[n']{\ms''}{W''}$ 
+    \end{itemize}
+    then we have that $(n', (\reg',\ms'')) \in \observations(W'')$
   \end{itemize}

### 2016-12-15 local-capability-machine.tex
changed configuration has stack to reg points to stack.
changed conclusion of scall lemma to take stack into account
typo in cnst region def
scall lemma: added unused to conclusion of Hyp-Cont
Updated scall works Hyp-Cont to work better

@@
   \iota^\cnst (v,n) &= (v,1,=,H^\cnst\;n)\\
-  H^\sta \; n \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \Rightarrow \forall a \in \dom(\ms)\ldotp \ms(a) = n \}
+  H^\cnst \; n \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \Rightarrow \forall a \in \dom(\ms)\ldotp \ms(a) = n \}
 \end{align*}
@@
 \begin{definition}
-  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ if
+  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ 
+  iff
   \begin{itemize}
@@
 \begin{definition}
-  $(\reg,\ms) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
-
+  We say that $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
+  iff
   \begin{itemize}
@@
     \overline{r_{\mathit{priv}}})) \text{ followed by } c_{\mathit{next}}$
-  \item $(\reg,\ms) \text{ has stack $\ms_\stk$ with $\ms_{\mathit{unused}}$ unused}$
+  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
   \item[Hyp-Callee] If
@@
     \item $\memSat[n-1]{\ms''}{W'}$
-    \item $(\reg',\ms'') \text{ has stack $\emptyset$ with
-        $\ms_{\mathit{unused}}'$ unused}$
+    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
     \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
@@
     \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)]$
+    \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
     \item for all $r$, we have that:
@@
           = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
-          \in \stdvr(W'') &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\}
+          \in \stdvr(W'') &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\} %revokeTemp here?
         \end{cases}
       \end{equation*}
-    \item $(\reg',\ms'')$ has stack $\ms_\stk$ with
-        $\ms_{\mathit{unused}}''$ unused for some $\ms_{\mathit{unused}}''$
-    \item $\memSat[n']{\ms''}{W''}$ 
-    \end{itemize}
-    then we have that $(n', (\reg',\ms'')) \in \observations(W'')$
+    \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
+    \end{itemize}
+\lau{Try to extract the ``private stack'' part, so it does not become a proof obligation in the proof that applies this lemma.}
+
+    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
   \end{itemize}
@@
   \begin{itemize}
-    \item $(\reg,\ms) \in \observations(W)$
+    \item $(\reg,\ms \uplus \ms_\stk \uplus \ms_{\var{unused}}) \in \observations(W)$
   \end{itemize}
@@
 \end{proof}
+
+\begin{lemma}[Execution and $\observations$]
+  \begin{align*}
+    & \forall n,i,\reg, \reg', \ms, \ms', \ms_r, W \ldotp\\
+    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \npair[n-i]{(\reg',\ms')} \in \observations(W) \\
+    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
+  \end{align*}
+
+\end{lemma}
 

### 2016-12-19 local-capability-machine.tex
updated name of observations anti reduction lemma
Comment for enter cap. Should they be global/local after jump?

@@
 \end{align*}
+\lau{At some point we talked about whether $\updatePcPerm{}$ should just give a $\glob$ capability all the time.
+I believe this is dangerous as it can be used to first get a $\glob$ read-capability for the stack and subsequently in a $\glob$ callback, you can start reading from someone elses private stack!}
 % TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
@@
 
-\begin{lemma}[Execution and $\observations$]
+\begin{lemma}[Anti-reduction for $\observations$]
   \begin{align*}

### 2016-12-21 local-capability-machine.tex
changed assumption in f3 lemma
indentation

@@
     \\ \lau{scall used to be dependent on linking, but now this part has been extracted to the fetch macro to make scall independent of linking.}
-%Load the entry $f$ from the linking table to a free register and jump to it.
+    % Load the entry $f$ from the linking table to a free register and jump to it.
   \item [Run the on stack restore code]
@@
 \subsubsection{Simple use of local capabilities}
-              The following programs only use local capabilities when they create protected return pointers.
-The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
+The following programs only use local capabilities when they create protected return pointers. The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
 
@@
 \begin{lemma}[Correctness lemma for \texttt{f2}]
-                \label{lem:correctness-f2}
-                let
-                \begin{align*}
-                  c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
-                  c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
-                  c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
-                  c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
-                  \reg & \in \Regs \\
-                  m & \defeq \hs_{f2} \uplus 
-                      \hs_\flag \uplus                
-                      \ms_{\var{link}} \uplus 
-                      \hs_\adv \uplus 
-                      \ms_{\malloc} \uplus 
-                      \ms_{\var{stk}} \uplus
-                      \ms_{\var{frame}} 
-                \end{align*}
-                where 
-                \begin{align*}
-                  &\dom(\hs_{f2}) = [\mathtt{f2}-\olf,\mathtt{2f}] \\
-                  &\dom(\hs_\flag) = [\flag,\flag] \\
-                  &\dom(\ms_\link) = [\link,\link+1]\\
-                  &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
-                  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-                  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
-                \end{align*}
-                and
-                \begin{itemize}
-                \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
-                \item $\ms_\flag = [\flag \mapsto 0]$
-                \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-                \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
-                \end{itemize}
-                if 
-                \[
-                  (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
-                \]
-                then
-                \[
-                  m'(\flag) = 0
-                \]  
-              \end{lemma}
-              \begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
-                Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
-                \begin{multline*}
-                  (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[k] \\
-                  (\reg_0\update{\pcreg}{\updatePcPerm{c_\adv}}\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''},m'')
-                \end{multline*}
-                Where 
-                \[
-                  m''  \defeq \hs_{f2} \uplus 
-                  \hs_\flag \uplus                
-                  \ms_\link \uplus 
-                  \hs_\adv \uplus 
-                  \ms_\malloc \uplus 
-                  \ms_\stk' \uplus
-                  \ms_{\var{frame}}
-                \]
-                Only the stack part of memory has changed. The stack and register-file at this point of execution is illustrated in figure~\ref{fig:f2-stack-before-adv}, where
-                \begin{itemize}
-                \item The range of authority for 
-                  \begin{itemize}
-                  \item $c_\stk'$ is $[\start_\stk,\addrend_\stk]$ (the entire stack)
-                  \item $c_\stk''$ is $[\start_\stk'',\addrend_\stk]$ (the unused part of the stack)
-                  \end{itemize}
-                  the two stack capabilities have the same permissions as $c_\stk$
-                \item $c_{f2}'$ points to the return address in f2's code.
-                \item $c_{\var{prp}}$ is a local enter-capability (with authority of all of the stack).
-                \item Notice: $f2$ has no return capability, so $r_0$ was not listed as a private register in the call.
-                \end{itemize}
-
-                \begin{figure}
-                  \centering
-                  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
-                    \multicolumn{2}{l}{Stack} \\
-                    \cline{2-2}
-                 &         \\
-                 & $\vdots$ \\
-                    \cline{2-2}
-                    $\start_\stk'':$   & $0$\\
-                    \cline{2-2}
-                    $c_\stk',c_\stk'' \rightarrow$   & $c_\stk'$ \\
-                    \cline{2-2}
-                 & $c_{f2}'$ \\
-                    \cline{2-2}
-                    $c_{\var{prp}} \rightarrow$   & Restore code \\
-                 & $\vdots$ \\
-                    \cline{2-2}
-                 & $1$\\
-                    \cline{2-2}
-                  \end{tabular}
-                  \hspace{1cm}
-                  \begin{tabular}{r |c |}
-                    \multicolumn{2}{r}{Register file} \\
-                    \cline{2-2}
-                    $\pcreg$ & $\updatePcPerm{c_\adv}$\\
-                    \cline{2-2}
-                    $r_0$  &  $c_{\var{prp}}$ \\
-                    \cline{2-2}
-                    $r_1$ & $c_\adv$ \\
-                    \cline{2-2}
-                    $r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
-                    \cline{2-2}
-                             & $0$ \\
-                    \cline{2-2}
-                             & $\vdots$ \\
-                    \cline{2-2}
-                  \end{tabular}
-                  \caption{Stack and register-file just after jump to $\adv$.}
-                  \label{fig:f2-stack-before-adv}
-                \end{figure}
-                Define 
-                \begin{align*}
-                  W = \;& [0 \mapsto \iota_{\malloc,0}] \\
-                        & [1 \mapsto \iota^\sta(\perma,\ms_{f2})]\\ 
-                        & [2 \mapsto \iota^\sta (\perma,\ms_\flag)]\\
-                        & [3 \mapsto \iota^\sta (\perma,\ms_\link)]\\
-                        & [4 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
-                        & [5 \mapsto \iota^\sta (\temp,\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})] \\
-                        & [6 \mapsto \iota^\pwl_{\start_\stk'', \addrend_\stk}]
-                \end{align*}
-                \lau{The interesting regions here are region 5 and 6.
-Region 5 governs our private part of the stack whereas region 6 governs the part of the stack we are about to release.
-I don't think it is strictly needed to make them $\temp$ for this example, but in general it is what we wish to do, so we can revoke them upon return.}
-                Show
-                \[
-                  \npair{(\start_\adv,\addrend_\adv)}\in\readCond{}(\glob)(\iota^\pwl_{\start_\adv,\addrend_\adv})
-                \]
-                Which amounts to
-                \[
-                  \iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}
-                \]
-                which is indeed the case by Lemma~\ref{lem:nwlp-subset-pwl}. Use this to apply the FTLR to get 
-                \begin{enumproof}
-                \item $\npair{c_\adv} \in \stder(W)$ \label{lem:f2-adv-er}
-                \end{enumproof}
-                Define
-                \[
-                  \ms  \defeq \hs_{f2} \uplus 
-                  \hs_\flag \uplus                
-                  \ms_\link \uplus 
-                  \hs_\adv \uplus 
-                  \ms_\malloc \uplus 
-                  \ms_\stk'
-                \]
-                ($m''$ without the frame) and show
-                \begin{enumproof}[resume]
-                \item $\heapSat[\ms]{n}{W}$ \label{lem:f2-mem-sat}\lau{The argument here is a bit informal}
-                  \begin{itemize}
-                  \item For $\hs_{f2}$, $\hs_\flag$, and $\ms_\link$ the regions $\iota^\sta(\perma,\ms_{f2})$,$\iota^\sta (\perma,\ms_\flag)$, and $\iota^\sta (\perma,\ms_\link)$ are designed to respectively accept each of these memory segments.
-                  \item $\hs_\adv$, showing this case is similar to item \ref{lem:f1-adv-mem-sat} in the proof of Lemma~\ref{lem:correctness-f1}.
-                  \item $\ms_\malloc$, this case follows from the specification of malloc and the assumption that $\iota_{\malloc,0}$ models $\ms_\malloc$.    
-                  \item $\ms_\stk'$, here we need to split the memory segment in two and for $n' < n$ argue
-                    \begin{itemize}
-                    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in
-                      H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})(W)$ \dominique{apply to world?}:\\ This region matches only this memory segment.
-                    \item case $\npair[n']{\ms_\stk'|_{[\start_\stk'', \addrend_\stk]}} \in H_{\start_\stk'', \addrend_\stk}(1)(W)$:\\ Show $\forall a \in [\start_\stk'', \addrend_\stk] \ldotp \ms_\stk'|_{[\start_\stk'', \addrend_\stk]}(a) \in \stdvr(W)$, trivially true since these memory cells contain 0.
-                    \end{itemize}
-                  \end{itemize}
-                \item $\npair{\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}} \in \stdrr(W)$ \label{lem:f2-reg-rel}
-                  \begin{enumproof}
-                  \item case $\npair{c_\stk''} \in \stdvr(W)$: Show: \label{f2:stack-in-vr}
-                    \begin{enumproof}
-                    \item $\npair{(\start_\stk'', \addrend_\stk)} \in \readCond{}(\local)(W)$ \\
-                      This amounts to showing 
-                      \[
-                        \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-                      \]
-                      which is indeed the case.
-                    \item $\npair{(\start_\stk'', \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$\\
-                      This amounts to showing 
-                      \[
-                        \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsupsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-                      \]
-                      which is indeed the case.
-                    \item $\npair{(\rwlx,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$ \label{f2:rwlx-exec-case} \\
-                      We will show this case in some detail, the others are similar. \\
-                      \dominique{Factor out a reusable lemma?}
-                      Let $n' < n$, $W' \futurewk W$, and $a \in [\start_\stk'', \addrend_\stk]$ be given. Show
-                      \[
-                        \npair[n']{((\rwlx,\local),\start_\stk'', \addrend_\stk,a)} \in \stder(W')
-                      \]
-                      If we can apply the FTLR at this point, then we are done. SFTS:
-                      \begin{enumproof}
-                      \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \readCond{}(\local)(W')$\\
-                        this amounts to showing
-                        \[
-                          \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-                        \]
-                        which is indeed the case.
-                      \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \writeCond{}(\local)(W')$ \\
-                        this amounts to showing
-                        \[
-                          \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsupsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-                        \]
-                        which is indeed the case.
-                      \end{enumproof}
-                    \item $\npair{(\rwx,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-                      Similar to case~\ref{f2:rwlx-exec-case}
-                    \item $\npair{(\exec,\start_\stk'', \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-                      Similar to case~\ref{f2:rwlx-exec-case}
-                    \end{enumproof}
-                  \item case $\npair{c_\adv} \in \stdvr(W)$:\\
-                    Follows from the FTLR.
-                  \item case $\npair{c_{\var{prp}}} \in \stdvr(W)$: \label{f2:c-prp}\\
-                    Let $n' < n$ and $W' \future W$ be given and show
-                    \[
-                      \npair[n']{\updatePcPerm{c_{\var{prp}}}} \in \stder(W').
-                    \]
-                    To this end let $n'' \leq n'$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n'']{\reg'} \in \stdrr$ be given and show
-                    \[
-                      \npair[n'']{(\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms')} \in \observations(W')
-                    \]
-                    Let $i \leq n''$, $m''$ and $\ms_f$ be given and assume
-                    \[
-                      (\reg'\update{\pcreg}{\updatePcPerm{c_{\var{prp}}}},\ms'\uplus\ms_f) \step[i]
-                      (\halted,m'')
-                    \]
-
-                    From $\heapSat[\ms']{n''}{W'}$ we know that that $\ms'$ contains memory segments $\hs_{f2}$, $\hs_\flag$, and $\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}$ (our local part of the stack).
-The restoration code remains unaltered, and so does the data on the stack, so the execution continues by popping the $c_\stk'$ to $r_\stk$ followed by a pop of $c_{f2}'$ to $\pcreg$.
-Continuing in $f2$, the restore code is popped off the stack which is the last of the $\mathtt{scall}$ macro-instruction.
-Now the top word of the stack is popped to $r_1$, and thanks to the world satisfaction, we know that this word is 1.
-The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.
-                    \dominique{Maybe you can factor out some sort of reusable antireduction lemma
-                      for this sort of argument?}
-
-                    All in all, there are no changes to the memory, so we can conclude
-                    \[
-                      m'' = \ms' \uplus \ms_f
-                    \]
-                    From Lemma~\ref{lem:priv-mono-like}, we can get $\ms''$ and $\ms_r$ such that
-                    \begin{itemize}
-                    \item $\ms' = \ms'' \uplus \ms_r$
-                    \item $\heapSat[\ms'']{n}{\revokeTemp{W}}$
-                    \end{itemize}
-                    Now downwards closure of memory segment satisfaction gives us the desired result.
-                  \item case $r\not\in \{\pcreg,r_0,r_1,r_\stk\}$: $\reg_0(r) = 0$. Trivial.
-                  \end{enumproof}
-                \end{enumproof}
-                From \ref{lem:f2-adv-er}, \ref{lem:f2-mem-sat}, and \ref{lem:f2-reg-rel} it follows that
-                \[
-                  \npair{(\reg_0\update{r_0}{c_{\var{prp}}}\update{r_1}{c_\adv}\update{r_\stk}{c_\stk''}\update{\pcreg}{c_\adv},\ms)} \in \observations(W)
-                \]
-                Now using $i$, the number of steps we need to halt, $\ms_f$ as the frame, $m'$ as the resulting memory and the assumption that the execution terminates successfully gives $W' \futurestr W$, $\ms'$, and $\ms_r$ such that
-                \begin{itemize}
-                \item $m' = \ms' \uplus\ms_r \uplus \ms_f$
-                \item $\heapSat[\ms']{n-i}{W'}$
-                \end{itemize}
-                The permanent region $\iota^\sta (\perma,\ms_\flag)$ ensures that $\ms_\flag$ is unchanged, so
-                \[
-                  m'(\flag) = 0
-                \]
-              \end{proof}
 
@@
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of code.
+  \item $\hs_\adv$ contains a read capability for $\hs_\link$ on its first address and otherwise it consists only of code.
   \end{itemize}
@@
 \end{lemma}
-
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
+\lau{proof using scall lemma}
+Assume the premises of the lemma.
+Now define 
+\[
+  W = [0 ]
+\]
+\end{proof}
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
+  \lau{This proof does not use scall lemma}
   This proof goes like the proof of Lemma~\ref{lem:correctness-f2} except for item \ref{f2:c-prp}
@@
       \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_]$ 
+      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
     \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$

### 2016-12-23 local-capability-machine.tex
small convenience change to scall
small scall change, f3 proof work in progress

@@
 \newcommand{\step}[1][]{\rightarrow_{#1}}
+
+% Conv defs
+\newcommand{\lookingat}[3]{\ensuremath{#1} \text{ is looking at } \ensuremath{#2} \text{ followed by } \ensuremath{#3}}
+\newcommand{\pointstostack}[3]{\ensuremath{#1} \text{ points to stack with } \ensuremath{#2} \text{ used and } \ensuremath{#3} \text{ unused}}
+
+% Macros
+\newcommand{\scall}[3]{\mathtt{scall} \; #1(#2,#3)}
 
@@
     c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link) \\
     \reg & \in \Regs \\
@@
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-  \item $\hs_\adv$ contains a read capability for $\hs_\link$ on its first address and otherwise it consists only of code.
+  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
   \end{itemize}
@@
 Now define 
-\[
-  W = [0 ]
-\]
+  \begin{align*}
+    W = \;& [0 \mapsto \iota_{\malloc,0}] \\
+          & [1 \mapsto \iota^\sta (\perma,\ms_\flag \uplus \ms_{f2})]\\
+          & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
+          & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
+  \end{align*}
+Further define
+\[
+\ms' = \ms_\flag \uplus \ms_{f2} \uplus \ms_\link \uplus \ms_\adv
+\]
+If we can show 
+\begin{equation}
+\label{eq:f3-obs1}
+  \npair[n+1]{(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms_{\malloc} \uplus \ms' \uplus \ms_{\adv} \uplus \ms_\stk)} \in \observations(W),
+\end{equation}
+then using $\ms_{\var{frame}}$ as the frame and $m'$ as the resulting memory, we get that $m' = \ms'' \uplus \ms_r \uplus \ms_{\var{frame}}$ for some $\ms'$ and $\ms_r$ s.t.
+$\heapSat[1]{\ms''}{W}$.
+Region $1$ guarantees that the flag is unchanged, so we have
+\[
+  m'(\flag) = 0
+\]
+So SFTS \ref{eq:f3-obs1}. To do so, we use Lemma \ref{lem:anti-red-obs}. By assumption, we have part of the execution, so we know
+\[
+  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_stk+1,\addrend_\stk} \uplus \ms_f),
+\]
+where
+\[
+  \lookingat{(\reg',\ms')}{\scall{r}{[]}{[]}}{c_{\var{next}}}
+\]
+where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction.
+Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.
+
+To show
+\[
+  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_stk+1,\addrend_\stk}} \in \observations(W)
+\]
+We use Lemma~\ref{lem:scall-works} which requires us to show
+\begin{enumproof}
+  \item $\memSat[n-i]{\ms'}{W}$ \\
+    Partition as follows:
+    \begin{enumproof}
+      \item $\ms_\malloc$: governed by $\iota_{\malloc,0}$, use malloc specification.
+      \item $\ms_\flag \uplus \ms_{f2}$: governed by island 1, only this memory segment is accepted.
+      \item $\ms_\link$: governed by island 2, only this memory segment is accepted.
+      \item $\ms_\adv$:\\
+        let $n' < n$ be given and show
+        \[
+          \npair[n']{\ms_\adv} \in H^\pwl_{\start_\adv,\addrend_\adv}\;1\;\xi^{-1}(W)
+        \]
+        this is trivial except for showing
+        \[
+          \npair[n'-1]{c_\link} \in \stdvr(W)
+        \]
+        amounts to showing
+        \[
+          \npair[n'-1]{\link,\link+1} \in \readCond{}(\glob)(W)
+        \]
+        which follows from
+        \begin{enumproof}
+          \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: \label{item:f3:adv1}\\
+            We will show
+            \begin{equation}
+              \label{eq:f3:adv-in-stder}
+              \forall W' \futurestr W \ldotp \npair{c_\adv} \in \stdvr(W')
+            \end{equation}
+            which will give us \ref{item:f3:adv1} using downwards closure.
+
+            Let $W' \futurestr W$ be given and show
+            \[
+              \npair{(\start_\malloc,\addrend_\malloc,\start_\malloc+\olf)} \in \entryCond{}(\glob)(W')
+            \]
+            to this end let $W'' \futurestr W'$ and $n' < n$ be given and show
+            \[
+              \npair[n']{\updatePcPerm{c_\adv}} \in \stder(W'')
+            \]
+            This follows from the FTLR if we can show
+            \[
+              \npair[n']{\start_\malloc,\addrend_\malloc} \in \readCond{}(\glob)(W'')
+            \]
+            $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}$ governs the adversary, so the result follows from Lemma~\ref{lem:nwlp-subset-pwl}.
+          \item $\npair[n'-1]{c_\malloc} \in \stdvr(W)$\\
+            Follows from the malloc specification.
+            \lau{I don't think this is actually the case at the moment. I guess we either need to show this based on the specification or just require it.}
+        \end{enumproof}
+    \end{enumproof}
+  \item Hyp-Callee\\
+    Assume
+    \begin{itemize}
+    \item $\ms_\stk |_{\start_stk+1,\addrend_\stk} = \ms_{\mathit{act}} \uplus \ms_{\mathit{unused}}$
+    \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$
+    \item $\memSat[n-1]{\ms''}{W'}$
+    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
+    \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
+    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
+      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
+      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
+    \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
+    \end{itemize}
+    where $\ms_{\mathit{act}}$ is the on stack activation code for this $\mathtt{scall}$.
+  \item Hyp-Cont\\
+    % TODO
+\end{enumproof}
+
 \end{proof}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
@@
 \begin{lemma}[$\mathtt{scall}$ works]
+  \label{lem:scall-works}
   If
@@
     \begin{itemize}
-    \item $\ms_{\mathit{unused}} = \ms_{\mathit{act}}
-      \uplus \ms_{\mathit{unused}}'$,
+    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{act}}
+      \uplus \ms_{\mathit{unused}}')$,
     \item $W' =
@@
 \begin{lemma}[Anti-reduction for $\observations$]
+\label{lem:anti-red-obs}
   \begin{align*}

### 2016-12-23 local-capability-machine.tex
added local requirement to points to definition
added lemma about stack capability in value relation, a bit more of f3 proof

@@
     \begin{itemize}
-    \item $\ms_\stk |_{\start_stk+1,\addrend_\stk} = \ms_{\mathit{act}} \uplus \ms_{\mathit{unused}}$
-    \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$
-    \item $\memSat[n-1]{\ms''}{W'}$
-    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
+    \item $\dom(\ms_\stk |_{\start_stk+1,\addrend_\stk}) = \dom(\ms_{\mathit{act}} \uplus \ms_{\mathit{unused}})$
+    \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,[\start_\stk \mapsto 1]\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$
+    \item $\memSat[n-i-1]{\ms''}{W'}$
+    \item $\reg'' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
     \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
-    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
-      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
-    \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
-    \end{itemize}
-    where $\ms_{\mathit{act}}$ is the on stack activation code for this $\mathtt{scall}$.
+    \item $\reg''= \reg_0[\pcreg\mapsto\updatePcPerm{\reg'(r)},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg'(r)]$ 
+    \item $\npair[n-i-1]{c_{\mathit{ret}}} \in \stdvr(W')$
+    \end{itemize}
+    for some $\ms_{\mathit{act}}$, $\ms_{\mathit{unused}}$, $\ms'$, $\reg''$, $c_{\mathit{ret}}$.
+
+    Using the FTLR, we get $\npair[n-i-1]{\updatePcPerm{c_\adv}} \in \stder(W')$, from
+    \begin{enumproof}
+      \item $\memSat[n-i-1]{\ms''}{W'}$ : By the above assumptions
+      \item $\npair[n-i-1]{\reg''} \in \stdvr(W')$:\\
+        show
+        \begin{enumproof}
+          \item $\npair[n-i-1]{c_{\mathit{ret}}} \in \stdvr(W')$ : by above assumptions.
+          \item $\npair[n-i-1]{reg''(r_\stk)} \in \stdvr(W')$ : by Lemma~\ref{lem:stack-cap-vr}.
+          \item $\npair[n-i-1]{c_\adv} \in \stdvr(W')$ : follows from \ref{eq:f3:adv-in-stder}.
+          \item The remaining registers we need to consider contain 0 and are thus trivial to show.
+        \end{enumproof}
+    \end{enumproof}
+    we get
+    \[
+      \npair[n-i-1]{(\ms'',\reg'')} \in \observations(W')
+    \]
   \item Hyp-Cont\\
@@
   \begin{itemize}
-  \item $\reg(r_\stk) =((\rwlx,\_),b_\stk,e_\stk,a_\stk)$
+  \item $\reg(r_\stk) =((\rwlx,\local),b_\stk,e_\stk,a_\stk)$
   \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
@@
   \end{align*}
-
 \end{lemma}
+
+\begin{lemma}[Stack capability in value relation]
+  \label{lem:stack-cap-vr}
+  \begin{align*}
+    & \pointstostack{\reg}{\ms'}{\ms} \land \\
+    & (\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))) \\
+    & \quad \Rightarrow \reg(r_\stk) \in \stdvr(W)
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stack-cap-vr}]
+  Say 
+\[
+  \reg(r_\stk) = c_\stk = ((\rwlx,\local),\start,\addrend,\_)
+\]
+  Show
+  \begin{enumproof}
+    \item $\npair{(\start,\addrend)} \in \readCond{}(\local)(W)$ : \label{pf:stack-cap-read}\\
+      Amounts to
+      \[
+        \iota^\pwl (\dom(\ms)) \nsubsim \iota^\pwl_{\start,\addrend}
+      \]
+      which is true as they are even equal.
+    \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W)$ : \label{pf:stack-cap-write}\\
+      Amounts to
+      \[
+        \iota^\pwl (\dom(\ms)) \nsupsim \iota^\pwl_{\start,\addrend}
+      \]
+      which is true as they are even equal.
+    \item $\npair{(\rwlx,\start,\addrend)} \in \execCond{}(\local)(W)$ : \label{pf:stack-cap-exec}\\
+      Let $n' < n$, $W' \futurewk W$, and $a\in \dom(\start,\addrend)$ be given and show
+      \[
+        \npair[n']{((\rwlx,\local),\start,\addrend,a)} \in \stder(W')
+      \]
+      to this end use the FTLR. To do so, we need to show
+      \begin{itemize}
+        \item $\npair[n']{(\start,\addrend)} \in \readCond{}(\local)(W')$
+        \item $\npair[n']{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W')$
+      \end{itemize}
+      As we are considering a public future world, some region still contains $\iota^\pwl (\dom(\ms))$, so it is the same argument as \ref{pf:stack-cap-read} and \ref{pf:stack-cap-write}, respectively.
+    \item $\npair{(\rwx,\start,\addrend)} \in \execCond{}(\local)(W)$ : \\
+      Same argument as \ref{pf:stack-cap-exec}, but also use \ref{lem:nwl-subset-pwl}.
+    \item $\npair{(\exec,\start,\addrend)} \in \execCond{}(\local)(W)$
+      Same argument as \ref{pf:stack-cap-exec}.
+  \end{enumproof}
+\end{proof}
 

### 2016-12-30 local-capability-machine.tex
Added region to past world in Hyp-Cont in scall lemma
changed Hyp-cont in scall to require obs rel wrt revoked world
First attempt at 'high-level' description of proof

@@
 \\ \lau{scall used to be dependent on linking, but now this part has been extracted to the fetch macro to make scall independent of linking.}
-    % Load the entry $f$ from the linking table to a free register and jump to it.
+%Load the entry $f$ from the linking table to a free register and jump to it.
                 \item [Run the on stack restore code]
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-f3} (high-level description)]
+Executing $f2$ until just after the jump in the first scall brings us to a configuration where the stack contains $1$ followed by some activation code followed by all zeros.
+The $\pcreg$-register contains an executable adversary capability, register $r_0$ contains a protected return pointer - that is a local enter capability for the execution code, and the $r_\stk$ contains a capability for the cleared part of the stack.
+
+At this point we can define a world with permanent regions
+\begin{itemize}
+\item fixing the assertion flag, the code of $f2$, and the linking table.
+\item the initial malloc region
+\item a $\iota^{\nwl,p}$ region
+\end{itemize}
+and temporary regions
+\begin{itemize}
+\item a region fixing the private part of the stack
+\item a $\iota^\pwl$ region for the rest of the stack 
+\end{itemize}
+From the FTLR, we get that in any future world of $W$, the adversary capability and its executable counter part is in the expression relation and thus safe to execute in suitable configurations.
+If the configuration we consider right now is suitable, then the execution produces a memory where the permanent invariants of $W$ are kept which means that the flag is $1$.
+
+To argue that the configuration is suitable, we need to argue that invoking the continuation produces an admissable result.
+As the continuation is a $\local$ capability, we take a public future world of $W$.
+In this public world, the private part of the stack remains the same as before the jump, so when we reach the assertion it succeeds and execution continues.
+At the point of the jump in the second scall, the stack contains $2$ instead of $1$, but otherwise essentially the same.
+Here we again use that it is safe to execute the adversary and that the continuation in this case halts immediately in a configuration where the assertion flag must be $1$.
+\end{proof}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
@@
 \end{proof}
-
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
@@
     \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)]$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unsused}}'')]$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
 
-    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
+    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(\revokeTemp{W''})$
   \end{itemize}
@@
   \begin{align*}
-    & \pointstostack{\reg}{\ms'}{\ms} \land \\
+    & \pointstostack{\reg}{\emptyset}{\ms} \land \\
     & (\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))) \\

### 2016-12-30 local-capability-machine.tex
f3 proof using scall

@@
 \[
-  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_stk+1,\addrend_\stk} \uplus \ms_f),
+  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_\stk+1,\addrend_\stk} \uplus \ms_f),
 \]
@@
 where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction.
-Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.
+Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_\stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.
 
@@
 \[
-  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_stk+1,\addrend_\stk}} \in \observations(W)
+  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_\stk+1,\addrend_\stk}} \in \observations(W)
 \]
@@
     \begin{itemize}
-    \item $\dom(\ms_\stk |_{\start_stk+1,\addrend_\stk}) = \dom(\ms_{\mathit{act}} \uplus \ms_{\mathit{unused}})$
+    \item $\dom(\ms_\stk |_{\start_\stk+1,\addrend_\stk}) = \dom(\ms_{\mathit{act}} \uplus \ms_{\mathit{unused}}')$
     \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,[\start_\stk \mapsto 1]\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$
@@
   \item Hyp-Cont\\
-    % TODO
-\end{enumproof}
-
+    Assume:
+    \begin{itemize}
+    \item $n' \leq n-i-2$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,[\start_\stk \mapsto 1])[\iota^\sta(\temp,\ms_{\mathit{unused}}'')]$
+    \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
+    \item for all $r$, we have that:
+      \begin{equation*}
+        \reg''(r)
+        \begin{cases}
+          = c_{\mathit{next}} &\text{ if } r = \pcreg\\
+          \in \stdvr(W'') &\text{ if $\reg''(r)$ is a global capability and } r \not\in \{\pcreg, r_\stk\}
+        \end{cases}
+      \end{equation*}
+    \item $\reg'' \text{ points to stack with $[\start_\stk \mapsto 1]$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
+    \end{itemize}
+    and show
+    \begin{enumproof}
+    \item $(\reg'',\ms'' \uplus [\start_\stk \mapsto 1] \uplus \ms_{\var{unused}}'') \in \observations(\revokeTemp{W''})$\\
+      As $W'' \futurestr W$, we know that the program, assertion flag, and linking table remain unchanged in $\ms''$.
+Given some frame $\ms_f'$, then the execution proceeds by first suceeding the assertion and then pushing 2 to the stack:
+      \[
+        (\reg'',\ms'' \uplus [\start_\stk \mapsto 1] \uplus \ms_{\var{unused}}'' \uplus \ms_f') \step[k] (\reg^{(3)},\ms'' \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}'' \uplus \ms_f')
+      \]
+      where 
+      \begin{itemize}
+      \item \lookingat{(\reg^{(3)},\ms'')}{\mathtt{scall}r([],[])}{c_{\var{next}}'}
+      \item \pointstostack{\reg^{(3)}}{[\start \mapsto 2]}{\ms_{\var{unused}}''}
+      \item $\reg^{(3)}(r) = c_\adv$
+      \end{itemize}
+      By Lemma~\ref{lem:anti-red-obs} it suffices to show
+      \begin{enumproof}
+      \item $\npair[n''-k]{(\reg^{(3)},\ms'' \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}'')} \in \observations(\revokeTemp{W''})$\\
+        Show this using Lemma~\ref{lem:scall-works}. So show:
+        \begin{enumproof}
+          \item $\memSat[n''-k]{\ms''}{\revokeTemp{W''}}$ is satisfied by one of the first Hyp-cont assumptions.
+          \item Hyp-Callee\\
+            Assume:
+            \begin{itemize}
+            \item $\dom(\ms_{\mathit{unused}}'') = \dom(\ms_{\mathit{act}}' \uplus \ms_{\mathit{unused}}^{(3)})$
+            \item $W^{(3)} = \revokeTemp{W''}[\iota^{\sta}(\temp,[\start_\stk \mapsto 2]\uplus\ms_{\mathit{act}}'),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}^{(3)}))]$
+            \item $\memSat[n''-k-1]{\ms^{(3)}}{W^{(3)}}$
+            \item $\reg^{(4)} \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}^{(3)}}$ unused}$
+            \item $\forall a \in \dom(\ms_{\var{unused}}^{(3)}) \ldotp \ms^{(3)}(a) = 0$
+            \item $\reg^{(4)}= \reg_0[\pcreg\mapsto\updatePcPerm{c_\adv},r_0 \mapsto c_{\mathit{ret}}', r_\stk \mapsto \_, r \mapsto c_\adv]$ 
+            \item $(n''-k-1,c_{\mathit{ret}}') \in \stdvr(W')$
+            \end{itemize}
+            This argument is almost identical to the one we just did for the first call:\\
+            Using the FTLR, we get $\npair[n-i-1]{\updatePcPerm{c_\adv}} \in \stder(W^{(3)})$. Which we use with
+            \begin{enumproof}
+              \item $\memSat[n''-k-1]{\ms^{(3)}}{W^{(3)}}$: By assumption.
+              \item $\npair[n''-k-1]{\reg^{(4)}}$: Show:
+                \begin{enumproof}
+                  \item $\npair[n''-k-1]{c_\adv} \in \stdvr(W^{(3)})$ by \ref{eq:f3:adv-in-stder}.
+                  \item $\npair[n''-k-1]{c_{\mathit{ret}}'}$ by assumption.
+                  \item $\npair[n''-k-1]{\reg^{(4)}(r_\stk)}$ by Lemma~\ref{lem:stack-cap-vr}.
+                \end{enumproof}
+            \end{enumproof}
+            to get
+            \[
+              \npair[n''-k-1]{(\reg^{(4)},\ms^{(3)})} \in \observations(W^{(3)})
+            \]
+          \item Hyp-Cont\\
+            Assume
+            \begin{itemize}
+            \item $n^{(3)} \leq n''-k-2$
+            \item $W^{(3)} \futurewk \revokeTemp{W''}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unused}}^{(3)})]$
+            \item $\memSat[n^{(3)}]{\ms^{(3)}}{\revokeTemp{W^{(3)}}}$ 
+            \item for all $r$, we have that:
+              \begin{equation*}
+                \reg^{(4)}(r)
+                \begin{cases}
+                  = c_{\mathit{next}} &\text{ if } r = \pcreg\\
+                  \in \stdvr(W'') &\text{ if $\reg^{(4)}(r)$ is a global capability and } r \not\in \{\pcreg, r_\stk\}
+                \end{cases}
+              \end{equation*}
+            \item $\reg' \text{ points to stack with $[\start_\stk \mapsto 2]$ used and $\ms_{\mathit{unused}}^{(3)}$ unused}$ for some $\ms_{\mathit{unused}}^{(3)}$
+            \end{itemize}
+            and show
+            \[
+              (n^{(3)}, (\reg^{(3)},\ms^{(3)} \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}^{(3)})) \in \observations(\revokeTemp{W^{(3)}})
+            \]
+            To this end let $\ms_f''$, $m''$, and $j \leq n^{(3)}$ be given and assume
+            \[
+              (\reg^{(3)},\ms^{(3)} \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}^{(3)} \uplus \ms_f'') \step[j] (\halted,m'')
+            \]
+            As the execution halts immediately, 
+            \[
+              m'' = \ms^{(3)} \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}^{(3)} \uplus \ms_f''
+            \]
+            By assumption we had $\memSat[n^{(3)}]{\ms^{(3)}}{\revokeTemp{W^{(3)}}}$ and the frame is unchanged, so we can split the memory as desired.
+        \end{enumproof}
+      \end{enumproof}
+    \end{enumproof}
+  \end{enumproof}
 \end{proof}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
@@
     \end{itemize}
+
     then we have that $(n-1,(\reg',\ms'')) \in \observations(W')$
@@
     \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unsused}}'')]$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unused}}'')]$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
       \end{equation*}
-    \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
-    \end{itemize}
-\lau{Try to extract the ``private stack'' part, so it does not become a proof obligation in the proof that applies this lemma.}
+    \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
+    \end{itemize}
 
@@
     & (\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))) \\
-    & \quad \Rightarrow \reg(r_\stk) \in \stdvr(W)
+    & \quad \Rightarrow \npair{\reg(r_\stk)} \in \stdvr(W)
   \end{align*}

### 2017-01-02 local-capability-machine.tex
Comment for points to stack
Description of f3 program and update to f3 proof

@@
 \subsection{Simple well-bracketedness/local capability example}
+\begin{comment}
 \begin{verbatim}
@@
 We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
-\lau{The above program and text are from before we introduced macros. They can be deleted when we have gotten used to the style of the new programs. }
+            \end{comment}
 
@@
 \end{verbatim}
+The assertion of $f3$ may seem a bit awkward because it is between two calls.
+If an adversary could capture the protected return pointer from the first call and save it until the second call, then the adversary could jump to it again.
+At this point the top of the stack would be 2, so when the execution reaches the assertion, it would fail.
+However, the producted return pointer is passed as a local capability, so the only place the adversary can store it is on the stack.
+The adversary loses control of the stack when control is returned to $f3$ where the $\mathtt{scall}$ makes sure to sanitise the stack and register file before control is passed back to the adversary.
+In other words, the adversary has no way to capture the continuation which makes the above safe and well-bracketed.
 
@@
 From the FTLR, we get that in any future world of $W$, the adversary capability and its executable counter part is in the expression relation and thus safe to execute in suitable configurations.
-If the configuration we consider right now is suitable, then the execution produces a memory where the permanent invariants of $W$ are kept which means that the flag is $1$.
+If the configuration we consider right now is suitable, then the execution produces a memory where the permanent invariants of $W$ are kept which means that the flag is $0$.
 
@@
 At the point of the jump in the second scall, the stack contains $2$ instead of $1$, but otherwise essentially the same.
-Here we again use that it is safe to execute the adversary and that the continuation in this case halts immediately in a configuration where the assertion flag must be $1$.
+Here we again use that it is safe to execute the adversary and that the continuation in this case halts immediately in a configuration where the assertion flag must be $0$.
 \end{proof}
@@
 $\heapSat[1]{\ms''}{W}$.
-Region $1$ guarantees that the flag is unchanged, so we have
+Region $1$ guarantees that the assertion flag is unchanged, so we have
 \[
@@
 \]
-So SFTS \ref{eq:f3-obs1}. To do so, we use Lemma \ref{lem:anti-red-obs}. By assumption, we have part of the execution, so we know
-\[
-  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_\stk+1,\addrend_\stk} \uplus \ms_f),
+So SFTS \ref{eq:f3-obs1}. To do so, we use Lemma \ref{lem:anti-red-obs}. Let $\ms_f$ be given. The execution proceeds as follows:
+\[
+  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms' \uplus \ms_\stk \uplus \ms_f) \step[i] (\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk|_{\start_\stk+1,\addrend_\stk} \uplus \ms_f),
 \]
@@
 \]
-where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction.
+where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction, namely \texttt{pop r1}.
 Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_\stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.
@@
 \[
-  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_\stk+1,\addrend_\stk}} \in \observations(W)
+  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_\stk+1,\addrend_\stk})} \in \observations(W)
 \]
@@
   \item $\memSat[n-i]{\ms'}{W}$ \\
-    Partition as follows:
+    Partition $\ms'$ as follows:
     \begin{enumproof}
@@
             \[
-              \npair{(\start_\malloc,\addrend_\malloc,\start_\malloc+\olf)} \in \entryCond{}(\glob)(W')
+              \npair{(\start_\adv,\addrend_\adv,\start_\adv+\olf)} \in \entryCond{}(\glob)(W')
             \]
@@
             \[
-              \npair[n']{\start_\malloc,\addrend_\malloc} \in \readCond{}(\glob)(W'')
+              \npair[n']{\start_\adv,\addrend_\adv} \in \readCond{}(\glob)(W'')
             \]
@@
             Follows from the malloc specification.
-            \lau{I don't think this is actually the case at the moment. I guess we either need to show this based on the specification or just require it.}
+            \lau{I don't think this is actually the case at the moment. I guess we either need to show this based on the specification or just require it in the specification.}
         \end{enumproof}
@@
     \end{itemize}
-    for some $\ms_{\mathit{act}}$, $\ms_{\mathit{unused}}$, $\ms'$, $\reg''$, $c_{\mathit{ret}}$.
+    for some $\ms_{\mathit{act}}$, $\ms_{\mathit{unused}}$, $\ms''$, $\reg''$, $c_{\mathit{ret}}$.
 
@@
       As $W'' \futurestr W$, we know that the program, assertion flag, and linking table remain unchanged in $\ms''$.
-Given some frame $\ms_f'$, then the execution proceeds by first suceeding the assertion and then pushing 2 to the stack:
+Given some frame $\ms_f'$, then the execution proceeds by first succeeding the assertion and then pushing 2 to the stack:
       \[
@@
       \item $\npair[n''-k]{(\reg^{(3)},\ms'' \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}'')} \in \observations(\revokeTemp{W''})$\\
-        Show this using Lemma~\ref{lem:scall-works}. So show:
+        Show this using Lemma~\ref{lem:scall-works}. Show:
         \begin{enumproof}
@@
             \item $\memSat[n''-k-1]{\ms^{(3)}}{W^{(3)}}$
-            \item $\reg^{(4)} \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}^{(3)}}$ unused}$
+            \item $\reg^{(4)} \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}^{(3)}$ unused}$
             \item $\forall a \in \dom(\ms_{\var{unused}}^{(3)}) \ldotp \ms^{(3)}(a) = 0$
@@
                 \begin{cases}
-                  = c_{\mathit{next}} &\text{ if } r = \pcreg\\
+                  = c_{\mathit{next}}' &\text{ if } r = \pcreg\\
                   \in \stdvr(W'') &\text{ if $\reg^{(4)}(r)$ is a global capability and } r \not\in \{\pcreg, r_\stk\}
@@
   \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
-  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$
+  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$ \lau{Maybe make it clear what happens when $\ms_\stk$ is empty}
   \item $b_\stk - 1\leq a_\stk$

### 2017-01-02 local-capability-machine.tex
Added another high-level proof of f3 - more true to the detailed proof.

@@
 
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-f3} (high-level description 2)]
+\lau{Another go at a proof sketch. More true to the proof below.}
+If we can show
+\begin{equation}
+  (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},\ms_{\malloc} \uplus \ms' \uplus \ms_{\adv} \uplus \ms_\stk) \in \observations(W),
+\end{equation}
+for a world $W$ where the assertion flag is permanently 0, then it is still 0 in any configuration the execution halts in.
+$W$ also needs to require the program and the linking table to permanently remain the same, have a region that governs $\malloc$ and a standard permanent no-write local region for the adversary.
+
+Due to Lemma~\ref{lem:scall-works} the \texttt{scall} lemma, for each \texttt{scall} we have to argue that the adversary and continuation produces results that respect the regions of $W$.
+Using Lemma~\ref{lem:anti-red-obs} the $\observations$ anti reduction lemma, it suffices to argue that each part of $f3$ between \texttt{scall}s produces admissable results.
+
+Executing until the first \texttt{scall} pushes 1 to the stack, so the invariants of $W$ are preserved.
+Due to the \texttt{scall} lemma, we need to argue that that the adversary and the continuation produce admissable results.
+
+Using the FTLR, we get that the executable capability for the adversary is in the $\stder$-relation.
+As we provide no arguments to the adversary, most of the conditions are satisfied by assumptions and Lemma~\ref{lem:stack-cap-vr}, which makes sure that the stack capability is in the value relation.
+
+With respect to the continuation, it is passed to the adversary as a local capability, so when we reason about it, we consider public future worlds.
+The \texttt{scall} uses temporary regions for the stack and these persist in public future worlds.
+This allows us to assume that the private part of the stack is the same after the adversary call, so it still contains 1.
+Further, the program, flags, and linking table remain the same in any kind of future world.
+Therefore, we know that the execution continues by poping 1 from the stack and then asserting that it is indeed 1.
+The assertion succeeds and 2 is pushed to the stack.
+At this point we reach another scall.
+No changes where made to the permanent part of the stack, so the invariants are still satisfied.
+At this point we use the \texttt{scall} lemma one last time.
+The adversary call code is well-behaved for the same reasons as in the first call.
+In the continuation, the execution halts immediately, so we use that the $\texttt{scall}$ lemma gives an assumption that the execution continues in a memory with all the invariants of $W$ preserved and so it also halts in such a memory.
+\end{proof}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]

### 2017-01-04 local-capability-machine.tex
Changed f3 scall high level proof a bit

@@
 
-Executing until the first \texttt{scall} pushes 1 to the stack, so the invariants of $W$ are preserved.
+Executing until the first \texttt{scall} only pushes 1 to the stack, so the invariants of $W$ are preserved.
 Due to the \texttt{scall} lemma, we need to argue that that the adversary and the continuation produce admissable results.
@@
 As we provide no arguments to the adversary, most of the conditions are satisfied by assumptions and Lemma~\ref{lem:stack-cap-vr}, which makes sure that the stack capability is in the value relation.
+Which gives us that the adversary produces an admissable result.
 
@@
 The \texttt{scall} uses temporary regions for the stack and these persist in public future worlds.
-This allows us to assume that the private part of the stack is the same after the adversary call, so it still contains 1.
+This allows us to assume that the private part of the stack still contains 1 after the call.
 Further, the program, flags, and linking table remain the same in any kind of future world.
-Therefore, we know that the execution continues by poping 1 from the stack and then asserting that it is indeed 1.
-The assertion succeeds and 2 is pushed to the stack.
+Therefore, we know that the execution continues by poping 1 from the stack and then asserting that it is indeed 1, which is indeed the case, so 2 is pushed to the stack.
 At this point we reach another scall.
@@
 The adversary call code is well-behaved for the same reasons as in the first call.
-In the continuation, the execution halts immediately, so we use that the $\texttt{scall}$ lemma gives an assumption that the execution continues in a memory with all the invariants of $W$ preserved and so it also halts in such a memory.
+The \texttt{scall} lemma lets us assume that the continuation continues in a memory that satisfies the invariants of $W$.
+The execution halts immediately in the continuation, so it produces an admissable result.
 \end{proof}

### 2017-01-05 local-capability-machine.tex
Changed the definition of worlds

@@
 \declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
+\declaretheorem[numbered=yes,name=Theorem,qed=$\blacksquare$]{theorem}
 \declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
@@
 \newcommand{\future}{\mathbin{\sqsupseteq}}
+\newcommand{\pub}{\var{pub}}
+\newcommand{\priv}{\var{priv}}
 \newcommand{\futurewk}{\mathbin{\sqsupseteq}^{\var{pub}}}
@@
 \newcommand{\Rel}{\plaindom{Rel}}
+\newcommand{\Rels}{\plaindom{Rels}}
 \newcommand{\States}{\plaindom{State}}
@@
   \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
     c_{g1} & \defeq ((\entry,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
@@
     c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link,\link) \\
     m & \defeq \hs_{g1} \uplus 
@@
   \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
-  \item $\hs_\adv$ contains a read-write capability for $\hs_\link$ on its first address and otherwise it consists only of instruction.
+  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
   \end{itemize}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using scall)]
+  Let $n$ be given and make the assumptions of the lemma. Define
+  \begin{align*}
+    W = & [0 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]\\
+        & [1 \mapsto \iota^\sta (\perma,\ms_\link)]\\
+        & [2 \mapsto \iota^\pwl_{\start_\stk, \addrend_\stk}]\\
+        & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
+        & [4 \mapsto \iota_{\malloc,0}]
+  \end{align*}
+  and
+  \[
+    \ms = \ms_{g1} \uplus 
+          \ms_\flag \uplus                
+          \ms_\link \uplus 
+          \ms_\adv \uplus 
+          \ms_\malloc 
+  \]
+  If we can show
+  \begin{equation}
+    \label{eq:f4:start-conf-in-obs}
+        \npair{(\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},\ms \uplus \ms_\stk)} \in \observations(W)
+  \end{equation}
+  then we are done. The termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 1 is permanent, so
+  \[
+    m(\flag) = 0
+  \]
+  So we need to show \ref{eq:f4:start-conf-in-obs}. 
+  To this end use the FTLR to show $\npair{c_\adv} \in \stder(W)$, so show
+  \begin{enumproof}
+    \item $\npair{(\start_\adv,\addrend_\adv)} \in \readCond{}(\glob)(W)$\\
+      %TODO
+    \item $\npair{(\start_\adv,\addrend_\adv)} \in \writeCond{}(\iota^\nwl,\glob)(W)$\\
+      %TODO
+  \end{enumproof}
+  in conclusion $\npair{c_\adv} \in \stder(W)$. We get \ref{eq:f4:start-conf-in-obs} if we show the following
+  \begin{enumproof}[resume]
+    \item $\memSat{\ms \uplus \ms_\stk}{W}$\\
+      %TODO
+    \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$\\
+      This amounts to showing
+      \begin{enumproof}
+        \item $\npair{c_\stk} \in \stdvr(W)$ \\
+          % TODO
+        \item $\npair{c_{g1}} \in \stdvr(W)$ \\
+          Let $n_1 < n$ and $W_1 \futurestr W$ and show
+          \[
+            \npair[n_1]{\updatePcPerm{c_{g1}}} \in \stder(W_1)
+          \]
+          To this end assume $n_2 \leq n_1$, $\memSat[n_2]{\ms_1}{W_1}$, and $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and show
+          \[
+            \npair[n']{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1} \in \observations(W_1)
+          \]
+          %TODO
+      \end{enumproof}
+  \end{enumproof}
+
+\end{proof}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-g1}]
@@
         & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
-        & [3 \mapsto \iota^\pwl]\\
+        & [3 \mapsto \iota^\pwl_{\start_\stk, \addrend_\stk}]\\ 
         & [4 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}] \\
@@
   \]
-  and use the FTLR on $c_\adv$ (possible due to lemma~\ref{lem:nwlp-subset-pwl} and the chose world). From this, we get
+  and use the FTLR on $c_\adv$ (possible due to lemma~\ref{lem:nwlp-subset-pwl} and the chosen world). From this, we get
   \[
@@
 \subsection{Worlds}
-Assume $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
+Assume a sufficiently large set of states $\States$ that at least contains the states used in this document.
+\begin{definition}
+  \[
+   \Rels = \{\Phi_\pub \times \Phi \in \powerset{\States^2}\times \powerset{\States^2} \mid \Phi_\pub, \phi \text{ is reflexive and transitive and } \Phi_\pub \subseteq \Phi \}
+  \]
+\end{definition}
+\begin{lemma}
+There exists $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
 \begin{align*}
   \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
-                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                                                & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{wk}}}) \\ \\
+                                                                & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
+                                                                & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_{\var{wk}})) \\ \\
   \xi_{\var{str}} :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
-                                                                & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                                                & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})))_{\future_{\var{str}}})
-\end{align*}
+                                                                & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
+                                                                & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_{\var{str}}))
+\end{align*}
+\end{lemma}
+
+\begin{theorem}
+  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are preordered \cofes{} and there exists an isomorphism $\xi$ such that
+  \begin{align*}
+    \xi : \Wor \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
+                                          & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
+                                          & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))))
+  \end{align*}
+ and for $W, W' \in \Wor$
+ \[
+   W \futurestr W' \Leftrightarrow \xi(W) \futurestr \xi(W')
+ \]
+and
+ \[
+   W \futurewk W' \Leftrightarrow \xi(W) \futurewk \xi(W')
+ \]
+\end{theorem}
+
 We now define the regions to be
@@
       &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi_{\var{wk}}(\hat{W}))
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
     \end{aligned}
@@
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xistr(\hat{W})})
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
     \end{aligned}

### 2017-01-05 local-capability-machine.tex
More world and region related changes

@@
   \[
-   \Rels = \{\Phi_\pub \times \Phi \in \powerset{\States^2}\times \powerset{\States^2} \mid \Phi_\pub, \phi \text{ is reflexive and transitive and } \Phi_\pub \subseteq \Phi \}
+   \Rels = \{\phi_\pub \times \phi \in \powerset{\States^2}\times \powerset{\States^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
   \]
 \end{definition}
+
+\begin{comment}
 \begin{lemma}
-There exists $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_{\var{wk}}$ and $\xi_{\var{str}}$ are isomorphisms such that
-\begin{align*}
-  \xi_{\var{wk}} :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
+There exists $\Worstr$ and $\Worwk$ are preordered \cofes{} and $\xi_\pub$ and $\xi_\priv$ are isomorphisms such that
+\begin{align*}
+  \xi_\pub :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
                                                                 & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                                                & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_{\var{wk}})) \\ \\
-  \xi_{\var{str}} :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
+                                                                & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_\priv)) \\ \\
+  \xi_\priv :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
                                                                 & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                                                & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_{\var{str}}))
+                                                                & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_\priv))
 \end{align*}
 \end{lemma}
+\end{comment}
 
@@
     \xi : \Wor \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
-                                          & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
-                                          & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))))
+                                          & \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
+                                          & \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))))
   \end{align*}
@@
   \Regions = & \{\revoked\} \uplus \\
-             & \{\temp\} \times \States \times \Rel \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) \uplus \\
-             & \{\perma\} \times \States \times \Rel \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments}))
-\end{align*}
+             & \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
+             & \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
+\end{align*}
+Let $\iota.v$ be the projection of the view of a region.
+
 And the worlds are
@@
 
-Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system:
+The two \emph{private future} region relations satisfies the following properties:
 \begin{mathpar}
-  \inferrule{  (s_1,s_2) \in \phi_2 \\
-    (\phi_1,H_1) = (\phi_2,H_2)}
-  {  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) }
+  \inferrule{  (s,s') \in \phi \\
+    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  {  (v,s',\phi_\pub',\phi',H') \futurestr (v',s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ }
+  { \revoked \futurestr \revoked }
+  \and
+  \inferrule{ }
+  { \revoked \futurestr (\temp,s,\phi_\pub,\phi,H) }
 \end{mathpar}
-our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
+The two \emph{public future} region relations satisfies the following properties:
 \begin{mathpar}
-  \inferrule{ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
-    v_1 = v_2 }
-  { (v_2,s_2,\phi_2,H_2) \futurewk (v_1,s_1,\phi_1,H_1) }
+  \inferrule{  (s,s') \in \phi_\pub \\
+    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  {  (v,s',\phi_\pub',\phi',H') \futurestr (v',s,\phi_\pub,\phi,H) }
   \and
@@
 \end{mathpar}
-For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
-\begin{mathpar}
-  \inferrule{ r_2 \futurewk  r_1 }
-  { r_2 \futurestr r_1 }
-  \and
-  \inferrule{ }
-  { \revoked \futurestr (\temp,s,\phi,H)}
-\end{mathpar}
-
-We now get two future world relations $\futurewk$ and $\futurestr$ are defined as one would expect.
-That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
+
+The two future world relations satisfy the following properties: They allow for any extension of the current world and all existing worlds are allowed to move to an appropriate future region. That is
 \begin{mathpar}
@@
         &\qquad \exists H,s \ldotp\\
-        &\qquad \quad W(r) = (\_,s,\_,H) \land \\
+        &\qquad \quad W(r) = (\_,s,\_,\_,H) \land \\
         &\qquad \quad \npair[n']{P(r)} \in H(s)(\xi^{-1}(W))\\
@@
   &\iota_{\start,\addrend}^\pwl : \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,H^\pwl_{\start,\addrend}) \\
+  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
   \\
-  &H^\pwl : \Addrs^2 \fun \States \fun (\Worwk \monnefun \UPred{\HeapSegments})\\
+  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
   &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
@@
                    \begin{cases}
-                     \revoked            & \text{if }W(r) = (\temp,s,\phi,H) \\
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
                      W(r)                & \text{otherwise}
@@
   &\iota_{\start,\addrend}^{\nwl} : \Regions \\
-  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,H^\nwl_{start,\addrend}) \\
+  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
   \\
   &\iota_{\start,\addrend}^{\nwl,p} : \Regions \\
-  &\iota_{\start,\addrend}^{\nwl,p} \defeq (\perma,1,=,H^\nwl_{start,\addrend})\\
+  &\iota_{\start,\addrend}^{\nwl,p} \defeq (\perma,1,=,=,H^\nwl_{start,\addrend})\\
   \\
-  &H^\nwl : \Addrs^2 \fun \States \fun (\Worstr \monnefun \UPred{\HeapSegments})\\
+  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
   &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
@@
 \begin{align*}
-  \iota^\sta (v,\ms) &= (v,1,=,H^\sta\;\ms)\\
+  \iota^\sta (v,\ms) &= (v,1,=,=,H^\sta\;\ms)\\
   H^\sta \; \ms \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
@@
 \begin{align*}
-  \iota^\cnst (v,n) &= (v,1,=,H^\cnst\;n)\\
+  \iota^\cnst (v,n) &= (v,1,=,=,H^\cnst\;n)\\
   H^\cnst \; n \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \Rightarrow \forall a \in \dom(\ms)\ldotp \ms(a) = n \}

### 2017-01-09 local-capability-machine.tex
scall lemma

@@
 \newcommand{\prp}{\var{prp}}
+\newcommand{\env}{\var{env}}
+\newcommand{\cls}{\var{cls}}
 
@@
 % Macros
-\newcommand{\scall}[3]{\mathtt{scall} \; #1(#2,#3)}
+\newcommand{\scall}[3]{\mathtt{scall} \; #1([#2],[#3])}
 
@@
 \[
-  \lookingat{(\reg',\ms')}{\scall{r}{[]}{[]}}{c_{\var{next}}}
+  \lookingat{(\reg',\ms')}{\scall{r}{}{}}{c_{\var{next}}}
 \]
@@
           \[
-            \npair[n']{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1} \in \observations(W_1)
-          \]
-          %TODO
+            \npair[n_2]{(\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1)} \in \observations(W_1)
+          \]
+          %TODO elaborate on what happens here using anti reduction lemma and malloc spec.
+          Use Lemma~\ref{lem:anti-red-obs}, so let $\ms_f'$ be given, then
+          \[
+            (\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1 \uplus \ms_f') \step[i] (\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls \uplus \ms_f')
+          \]
+          for some $i$. Where
+          \begin{itemize}
+          \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{4}{\iota_\malloc}} $
+          \item $\iota_\malloc \futurewk \iota_{\malloc,0}$
+          \item $\ms_x = [x \mapsto 0]$
+          \item $c_x = ((\rwx,\glob),x,x,x)$
+          \item $\ms_\env = [\env \mapsto c_x]$
+          \item $c_\env = ((\rwx,\glob,\env,\env,\env))$
+          \item $c_{f4}$ is $c_{g1}$ updated to point at $f4$.
+          \item $\ms_\cls$ contains $c_\env$, the $c_{f4}$, and the activation code.
+          \item $c_\cls$ is a global enter capability for $\ms_\cls$ pointing to the activation code.
+          \item $\reg_2 = reg_0\update{\pcreg}{\updatePcPerm{\reg_1(r_0)}}\update{r_0}{\reg_1(r_0)}\update{r_1}{c_\cls}$
+          \end{itemize}
+          For world
+          \[
+            W_2 = W_1\update{4}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\iota{i_2}{\iota_x}
+          \]
+          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is defined as
+          \begin{align*}
+            \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
+            \phi_\pub & = \{(0,1)\}^* \\
+            \phi      & = (1,0) \union \phi_\pub \\
+            H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \uplus \{\npair[0]{\ms}\}
+          \end{align*}
+          % TODO show $\iota_x$ is a region.
+          show
+          \begin{equation}
+            \label{eq:g1-obs-to-show}
+            \npair[n_2-i]{(\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls)} \in \observations(W_2)
+          \end{equation}
+          If $\reg_1(r_0).\perm \not\in \{\entry, \exec, \rwx, \rwlx\}$, then the execution fails after the jump and \label{eq:g1-obs-to-show} is thus trivially true.
+
+          If $\reg_1(r_0).\perm \in \{\entry, \exec, \rwx, \rwlx\}$, then either $\execCond{}$ or $\entryCond{}$ holds for the capability in $\reg_1(r_0)$.
+Now use $W_2 \futurewk W_1$ with the approriate condition to get
+          \[
+            \npair[n_2-i]{\updatePcPerm{\reg_1(r_0)}} \in \stder(W_2)
+          \]
+          which in turn gives us \ref{eq:g1-obs-to-show} if we can show the following
+          \begin{enumproof}
+            \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls}{W_2}$
+              % TODO follows from \ms_1 heap sat and the new regions.
+            \item $\npair[n_2-i]{\reg_2} \in \stdrr$ \\
+              Amounts to showing
+              \begin{enumproof}
+                \item $\npair[n_2-i]{\reg_2(r_0)} \in \stdvr(W_2)$ by assumption $\npair[n_2]{\reg_1} \in \stdrr(W_1)$ and $\stdvr$ monotonicity wrt. $\futurewk$
+                \item $\npair[n_2-i]{c_\cls} \in \stdvr(W_2)$ \\
+                  Let $n_3 < n_2 - i$ and $W_3 \futurestr W_2$ be given and show
+                  \[
+                    \npair[n_3]{\updatePcPerm{c_\cls}} \in \stder(W_3)
+                  \]
+                  To this and let $n_4 \leq n_3$, $\memSat[n_4]{\ms_3}{W_3}$, and $\npair[n_4]{\reg_3} \in \stdrr(W_3)$ and show
+                  \begin{equation}
+                    \label{eq:g1-c-cls-obs}
+                    \npair[n_4]{(reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3)} \in \observations(W_3)
+                  \end{equation}
+                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W}_3}$ (using Lemma~\ref{lem:priv-mono-like2})
+
+                  Use Lemma~\ref{lem:anti-red-obs}, let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern.
+By the operational semantics, we know\footnote{the execution may fail, but then the confifuration is trivially in the observation relation.}
+                  \[
+                    (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f'') \step[j] (\reg_4,\ms_4 \uplus \ms_3^t \uplus \ms_f'')
+                  \]
+                  where
+                  \begin{itemize}
+                  \item \lookingat{reg_4,\ms_4}{\scall{r_1}{}{r_0,r_1,r_\env}}{c_{\var{next}}}
+                    \begin{itemize}
+                    \item $c_{\var{next}}$ is the capability pointing to the next instruction.
+                    \end{itemize}
+                  \item \pointstostack{\reg_4}{\emptyset}{\ms_{\var{unused}}}
+                    \begin{itemize}
+                    \item \texttt{prepstack} did not fail, so the capability must be $\rwlx$ and follow the stack convention.
+                    \end{itemize}
+                  \item $\reg_3(r_0)$ is a $\glob$ capability.
+                    \begin{itemize}
+                    \item \texttt{reqglob} did not fail
+                    \end{itemize}
+                  \item $\ms_4(x) = 0$
+                  \item $\reg_3(r_\env) = c_\env$
+                  \end{itemize}
+                  region $i_2$ (the $\iota_x$ region) can be in either state $0$ or $1$, so to make sure it is in state $0$, we use a private transition.
+So let $W_4$ be $\revokeTemp{W_3}$ with region $i_2$ in state $0$.
+              \end{enumproof}
+          \end{enumproof}
       \end{enumproof}
@@
   \begin{itemize}
-  \item $\heapSat[\ms]{n}{W}$ 
+  \item $\heapSat[\ms]{n}{\revokeTemp{W}}$ 
   \item $(\reg,\ms) \text{ is looking at }
@@
 
-    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(\revokeTemp{W''})$
+    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
   \end{itemize}
@@
   \begin{align*}
-    & \forall n,i,\reg, \reg', \ms, \ms', \ms_r, W \ldotp\\
-    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \npair[n-i]{(\reg',\ms')} \in \observations(W) \\
+    & \forall n,i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
+    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \npair[n-i]{(\reg',\ms')} \in \observations(W') \land W' \futurestr W \\
     & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
@@
 \lau{$a \leq b$ that the interpretation is not empty.}
-
-\begin{lemma}
-  Say the instructions for \texttt{scall} $r([r_{j_1},\dots,r_{j_n}],[r_{k_1},\dots, r_{k_m}])$ are $i_0,\dots,i_l$, then
-  \begin{align*}
-    & m = \ms \uplus \ms_\stk \uplus \ms_f \\
-    & \memSat{\ms \uplus \ms_\stk}{W}
-    & \npair{\reg[\pc \mapsto c_{\var{scall}}][r_\stk \mapsto c_\stk]} \in \stdrr(W) \\
-    & \ms_\stk = m | _{\start_\stk,\addrend_\stk} \\
-    & c_\stk = ((\rwlx,\gl),\start_\stk,\addrend_\stk,\addr) \\
-    & \start_\stk-1 \leq \addr \land \start_\stk \leq \addrend_\stk \land \addr \leq \addrend_\stk \\
-    & c_{\var{scall}} = ((\perm,\gl)\start,\addrend,\addr) \\
-    & \perm \in \{\exec, \rwx, \rwlx\} \\
-    & \addrend - \addr \leq l  \\
-    & \ms(\addr) = i_0, \dots, \ms(\addr+l) = \mathtt{jmp r}
-    & (\reg[\pc \mapsto c_{\var{scall}}][r_\stk \mapsto c_\stk],m) \step[i'] (\reg_0[r_{k_1} \mapsto \reg(r_{k_1})]\dots[r_{k_m} \mapsto \reg(r_{k_m})][\pc \mapsto c_{\var{scall}}'][r_\stk \mapsto c_\stk'],m) \\
-    & c_{\var{scall}}'.\mathrm{a} = \addr+l \\
-    & \quad \Rightarrow \\
-    & \qquad \exists \ms_\stk', \ms', \ms_r \ldotp \\
-    & \qquad \quad W' = \revokeTemp{W}[\var{rn}_1\mapsto\iota^\sta(\temp, )][\var{rn}_2 \mapsto \iota^\pwl] \\
-    & \qquad \quad \memSat[n-i']{\ms_\stk'\uplus \ms'}{W'}
-  \end{align*}
-\end{lemma}
 

### 2017-01-10 local-capability-machine.tex
Added FTLR
more of the g1 proof, missing last cont argument

@@
 \newcommand{\cls}{\var{cls}}
+\newcommand{\unused}{\var{unused}}
 
@@
       \begin{itemize}
-      \item \lookingat{(\reg^{(3)},\ms'')}{\mathtt{scall}r([],[])}{c_{\var{next}}'}
+      \item \lookingat{(\reg^{(3)},\ms'')}{\mathtt{scall}\;r([],[])}{c_{\var{next}}'}
       \item \pointstostack{\reg^{(3)}}{[\start \mapsto 2]}{\ms_{\var{unused}}''}
@@
               % TODO follows from \ms_1 heap sat and the new regions.
-            \item $\npair[n_2-i]{\reg_2} \in \stdrr$ \\
+            \item $\npair[n_2-i]{\reg_2} \in \stdrr(W_2)$ \\
               Amounts to showing
@@
                   \end{equation}
-                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W}_3}$ (using Lemma~\ref{lem:priv-mono-like2})
+                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W_3}}$ (using Lemma~\ref{lem:priv-mono-like2})
 
@@
                   \begin{itemize}
-                  \item \lookingat{reg_4,\ms_4}{\scall{r_1}{}{r_0,r_1,r_\env}}{c_{\var{next}}}
+                  \item \lookingat{(reg_4,\ms_4)}{\scall{r_1}{}{r_0,r_1,r_\env}}{c_{\var{next}}}
                     \begin{itemize}
@@
                     \end{itemize}
-                  \item $\reg_3(r_0)$ is a $\glob$ capability.
+                  \item $\reg_3(r_1)$ is a $\glob$ capability.
                     \begin{itemize}
@@
 So let $W_4$ be $\revokeTemp{W_3}$ with region $i_2$ in state $0$.
-              \end{enumproof}
-          \end{enumproof}
+We then have
+                  \[
+                    \memSat[n_4-j]{\ms_4}{W_4}
+                  \]
+                  Now we can use Lemma~\ref{lem:scall-works} to show:
+                  \[
+                    \npair[n_4-j]{(\reg_4,\ms_4 \uplus \ms_3^t)} \in \observations(W_4)
+                  \]
+                  \begin{enumproof}
+                    \item $\memSat[n_4-j]{\ms_4}{\revokeTemp{W_4}}$: follows from $W_4 = \revokeTemp{W_4}$
+                    \item Hyp-Callee\\
+                      We know $\npair[n_4]{\reg_3(r_1)} \in \stdvr(W_3)$.
+If this is not a capability that becomes executable when jumped to, then the execution fails, so the register memory segment pair is trivially in the observation relation.
+If it is executable, then either the $\execCond{}$ or the $\entryCond{}$ holds for appropriate values.
+We also know that it is a global capability, so we can use it with private future worlds.
+We have $W_5 = \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset)][\iota^\pwl(\dom(\ms_\unused))] \futurestr W_3$, so by the conditions, we have
+                      \[
+                        \npair[n_4-j]{\updatePcPerm{reg_3(r_1)}} \in \observations(W_5)
+                      \]
+                      Now it suffices to show
+                      \begin{enumproof}
+                        \item $\memSat[n_4-j-1]{\ms_5}{W_5}$ for some $\ms_5$ which is one of the assumptions of Hyp-Callee.
+                        \item $\npair[n_4-j-1]{\reg_5} \in \stdrr(W_5)$ where $\reg_5$ is as described in the \texttt{scall} lemma Hyp-callee premise.\\
+                          Amounts to showing:\lau{too deply nested for environments}\\
+                          1) $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.
+
+                          2) The protected return pointer is in the value relation by Hyp-callee assumption, and the stack capability is handled by lemma~\ref{lem:stack-cap-vr}.
+                      \end{enumproof}
+                      which gives us
+                      \[
+                        \npair[n_4-j-1]{(\reg_5,\ms_5)} \in \observations(W_5)
+                      \]
+                    \item Hyp-Cont\\
+                      Assume
+                      \begin{itemize}
+                      \item $n_5 \leq n_4-j-2$
+                      \item $W_6 \futurewk W_5$ (as defined in the Hyp-Callee case).
+                      \item $\memSat[n_4]{\ms_6}{\revokeTemp{W_6}}$
+                      \item $\reg_6(\pcreg) = c_{\var{next}}$, $\reg_6(r_0) = \reg_3(r_0)$, $\reg_6(r_1) = \reg_3(r_1)$, $\reg(r_\env) = c_\env$
+                      \item \pointstostack{\reg_6}{\emptyset}{\ms_\unused'}
+                      \end{itemize}
+                      Use the $\observations$-anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) followed by the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
+Given $\ms_f'''$, we know by the operational semantics and the fact that the program hasn't changed that
+                      \[
+                        (\reg_6,\ms_6 \uplus \ms_\unused''' \uplus \ms_f''') \step[k] (\reg_7,\ms_6\update{x}{1} \uplus \ms_\unused'' \uplus \ms_f''')
+                      \]
+                      where
+                      \begin{itemize}
+                      \item \lookingat{(\reg_7,\ms_6\update{x}{1})}{\mathtt{scall}\;r([],[r_0,r_\env])}{c_{\var{next}}'}\\
+                        $c_{\var{next}}'$ is the current $\pcreg$ capability but looking at \texttt{load} $r_1$ $x$.
+                      \item $\reg_7(r_0,r_1,r_\env,r_\stk) = \reg_6(r_0,r_1,r_\env,r_\stk)$
+                      \end{itemize}
+                      In $\revokeTemp{W_6}$, we don't know which state the $\iota_x$ region is in, but state $1$ is reachable via a public transition, so let $W_7$ be $\revokeTemp{W_6}$ with region $i_2$ in state $1$.
+It follows easily that
+                      \[
+                        \memSat[n_4-k]{\ms_6\update{x}{1}}{W_7}
+                      \]
+                      We continue the proof in item \ref{item:proof-cont}
+                  \end{enumproof}
+              \end{enumproof}
+          \end{enumproof}
+      \end{enumproof}
+      \item \label{item:proof-cont} At this point, we apply the \texttt{scall} lemma, to get
+        \[
+          \npair[n_4-k]{(\reg_7,\ms_6\update{x}{1} \uplus \ms_\unused'')} \in \observations(W_7)
+        \]
+        show
+        \begin{enumproof}
+          \item $\memSat[n_4-k]{\ms_6\update{x}{1}}{\revokeTemp{W_7}}$, follows from $W_7= \revokeTemp{W_7}$.
+          \item Hyp-Callee
+            % Use that it is the same $r_1$
+            % TODO
+          \item Hyp-Cont
+
         \end{enumproof}
@@
 
+\begin{theorem}[Fundamental theorem of logical relations]
+\lau{do we want the FTLR to also talk about everything that is trivially in the FTLR.}
+  \label{thm:ftlr}
+  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
+  if
+  \[
+    \begin{gathered}
+      \perm = \exec \land\\
+      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+    \end{gathered}
+  \]
+  or
+  \[
+    \begin{gathered}
+      \perm = \rwx \land \\ 
+      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
+      \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
+    \end{gathered}
+  \]
+  or
+  \[
+    \begin{gathered}
+      \perm = \rwlx \land\\ 
+      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\ 
+      \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
+    \end{gathered}
+  \]
+  then
+  \[
+    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
+  \]
+\end{theorem}
+
 \begin{definition}
@@
 \begin{lemma}[Global capabilities monotone wrt $\futurestr$]
+  \label{lem:stdvr-glob-priv-mono}
   \begin{align*}

### 2017-01-16 local-capability-machine.tex
Fixed op sem bug permitting lea on enter cap.
Added a figure for the worlds
test
Fixed disjoint mem sat lemma
Naive attempt at adding framing to scall lemma
small example w/ same problem as akward example

@@
                                                  \quad\text{and $\memreg(r_1) = \stdcap$}\\
+                                                 \quad\text{and $\perm \neq \entry$}\\
                                                  \quad\text{and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$}
@@
 
+\subsection{Inverted control and return from closure}
+If we can prove the below example using the scall lemma, then we should also be able to prove the akward example.
+\begin{lstlisting}
+g2:  move $\pcreg$ $r_3$
+     lea $r_3$ $\dots$
+     crtcls $[]$ $r_3$
+     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
+2g:  jmp $r_0$
+f5:  reqglob $r1$
+     prepstack $r_\stk$
+     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
+     mclear $r_\stk$
+     rclear $\RegName \setminus \{r_0,\pcreg\}$
+5f:  jmp $r_0$
+\end{lstlisting}
+
 \subsection{Variant of the ``awkward'' example}
@@
           \[
-            W_2 = W_1\update{4}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\iota{i_2}{\iota_x}
+            W_2 = W_1\update{4}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\update{i_2}{\iota_x}
           \]
@@
           \item Hyp-Cont
-
+            % From c_stk in val rel of W_3 get that W_3 must have iota pwl region.
+            % 
         \end{enumproof}
@@
 
-  Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
-  words for stack pointer and return pointer and some fixed number of words for
-  the standard return instructions).
+  % Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
+  % words for stack pointer and return pointer and some fixed number of words for
+  % the standard return instructions).
 \end{lemma}
@@
   \begin{align*}
-    & \forall n \ldotp \forall \ms, \ms' \ldotp \forall W, W' \ldotp\\
-    & \quad \memSat{\ms}{W} \land \memSat{\ms'}{W'} \Rightarrow \\
+    & \forall n \ldotp \forall \ms, \ms', \ms'' \ldotp \forall W, W', W'' \ldotp\\
+    & \quad \ms'' = \ms \uplus \ms' \land W'' = W \uplus W' \land \memSat{\ms}{W} \land \memSat{\ms'}{W'} \Rightarrow \\
     & \qquad \heapSat[\ms \uplus \ms' ]{n}{W \uplus W'}

### 2017-01-23 local-capability-machine.tex
changed revoke to remember state
typo in future world rel
More changes related to reinstating revoked regions
Allowed revoking/reinstating to also take transition in state machine (for transitivity).
A few minor changes + notes.

--- 2017-01-10 local-capability-machine.tex
@@
 \usepackage{listings}
+\usepackage{hyperref}
 \lstset{
@@
 \end{flushright}
+
+\tableofcontents
+
 \section{Domains and Notation}
@@
               \end{lemma}
-
               \begin{proof}[Proof of Lemma~\ref{lem:correctness-f1}]
@@
 
-\section{Logical Relation'}
+\section{Logical Relation}
 \subsection{Worlds}
@@
 \begin{align*}
-  \xi_\pub :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
+  \xi_\pub :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
                                                                 & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
                                                                 & \{\perma\} \times \States \times \Rels \times (\States \fun (\Worstr \monnefun \UPred{\HeapSegments})),\future_\priv)) \\ \\
-  \xi_\priv :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
+  \xi_\priv :   \Worstr \cong \blater (\nats \finparfun ( & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
                                                                 & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
@@
   \begin{align*}
-    \xi : \Wor \cong \blater (\nats \finparfun ( & \{\revoked\} + \\
+    \xi : \Wor \cong \blater (\nats \finparfun ( & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
                                           & \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
@@
 \begin{align*}
-  \Regions = & \{\revoked\} \uplus \\
+  \Regions = & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
              & \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
@@
 \]
-Where $\RegionNames = \nats$.
+where $\RegionNames = \nats$.
 
@@
   \inferrule{ }
-  { \revoked \futurestr \revoked }
+  { (\revoked,s,\phi_\pub,\phi,H) \futurestr (\revoked,s,\phi_\pub,\phi,H) }
   \and
-  \inferrule{ }
-  { \revoked \futurestr (\temp,s,\phi_\pub,\phi,H) }
+  \inferrule{ (s,s') \in \phi }
+  { (\revoked,s',\phi_\pub,\phi,H) \futurestr (\temp,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ (s,s') \in \phi }
+  { (\temp,s',\phi_\pub,\phi,H) \futurestr (\revoked,s,\phi_\pub,\phi,H) }
 \end{mathpar}
@@
     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-  {  (v,s',\phi_\pub',\phi',H') \futurestr (v',s,\phi_\pub,\phi,H) }
+  {  (v,s',\phi_\pub',\phi',H') \futurewk (v',s,\phi_\pub,\phi,H) }
   \and
   \inferrule{ }
-  { \revoked \futurewk \revoked }
+  { (\revoked,s,\phi_\pub,\phi,H) \futurewk (\revoked,s,\phi_\pub,\phi,H) }
 \end{mathpar}
@@
 
-Define the function $\active{}$ as follows
+Define the function $\activeReg{\cdot}$ as follows
 \begin{align*}
@@
 \lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
+\dominique{The comment immediately above is outdated, right?}
 Revoking all temporary regions:
@@
                    \begin{cases}
-                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
+                     (\revoked,s,\phi_\pub,\phi,H)            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
                      W(r)                & \text{otherwise}
@@
   \begin{itemize}
-  \item $\heapSat[\ms]{n}{\revokeTemp{W}}$ 
+  \item $\memSat[n]{\ms}{\revokeTemp{W}}$ 
+  \item $\dom(\ms_f) \cup (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
   \item $(\reg,\ms) \text{ is looking at }
@@
     \item $W' =
-      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
+      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}')),\iota^\sta (\temp,\ms_f)]$,
     \item $\memSat[n-1]{\ms''}{W'}$
@@
     \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unused}}'')]$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unused}}'')][\iota^\sta(\temp,\ms_f)]$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
 
-    then we have that $(n', (\reg',\ms'' \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
+    then we have that $(n', (\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
   \end{itemize}
@@
   \begin{itemize}
-    \item $(\reg,\ms \uplus \ms_\stk \uplus \ms_{\var{unused}}) \in \observations(W)$
-  \end{itemize}
-
-  Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
-  words for stack pointer and return pointer and some fixed number of words for
-  the standard return instructions).
+    \item $(\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}) \in \observations(W)$
+  \end{itemize}
+
+  % Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
+  % words for stack pointer and return pointer and some fixed number of words for
+  % the standard return instructions).
 \end{lemma}
@@
   \end{align*}
+  \dominique{The above still needs updating now that revoked regions remember
+    what they used to look like.}
   For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$, we have $W' \futurestr W$. Now define the following memory segment:

### 2017-01-24 local-capability-machine.tex
Fix typo's.

@@
                 \item[Store register contents to environment]
-                \item[Allocate memory for record with environment capability, code capability, and acitvation code]
+                \item[Allocate memory for record with environment capability, code capability, and activation code]
                 \item[Store capabilities and activation code to record]
-                \item[Restrict the capability for the ``clojure pair'' to an enter capability]
+                \item[Restrict the capability for the ``closure pair'' to an enter capability]
                 \item[Activation code:] \forcenewline
@@
                 \item[Store register contents to environment] Store the contents of each of the registers $r_1,\dots,r_n$ to the newly allocated memory.
-                \item[Allocate memory for record with environment capability, code capability, and acitvation code] Allocate a new piece of memory with room for a capability for the environment.
+                \item[Allocate memory for record with environment capability, code capability, and activation code] Allocate a new piece of memory with room for a capability for the environment.
                 \item[Store capabilities and activation code to record] Store the environment capability and code capability in the record followed by the activation code. 
-                \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the acitvation code and restrict it to a global enter-capability.
+                \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the activation code and restrict it to a global enter-capability.
                 \item[Activation code:] \forcenewline
@@
   \item $\dom(\ms_f) \cup (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
+    \dominique{you mean $\cap$, right?}
   \item $(\reg,\ms) \text{ is looking at }

### 2017-01-24 local-capability-machine.tex
more added to awkward proof

--- 2017-01-23 local-capability-machine.tex
@@
 \usepackage{listings}
-\usepackage{hyperref}
 \lstset{
@@
 \newcommand{\unused}{\var{unused}}
+\newcommand{\act}{\var{act}}
 
@@
 \end{flushright}
-
-\tableofcontents
-
 \section{Domains and Notation}
@@
               \end{lemma}
+
               \begin{proof}[Proof of Lemma~\ref{lem:correctness-f1}]
@@
                   Use Lemma~\ref{lem:anti-red-obs}, let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern.
-By the operational semantics, we know\footnote{the execution may fail, but then the confifuration is trivially in the observation relation.}
+By the operational semantics, we know\footnote{the execution may fail, but then the configuration is trivially in the observation relation.}
                   \[
@@
                   \[
-                    \npair[n_4-j]{(\reg_4,\ms_4 \uplus \ms_3^t)} \in \observations(W_4)
-                  \]
+                    \npair[n_4-j]{(\reg_4,\ms_4 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused)} \in \observations(W_4)
+                  \]
+                  where $\ms_r$ is the local frame of the \texttt{scall} lemma.
                   \begin{enumproof}
@@
 We also know that it is a global capability, so we can use it with private future worlds.
-We have $W_5 = \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset)][\iota^\pwl(\dom(\ms_\unused))] \futurestr W_3$, so by the conditions, we have
-                      \[
-                        \npair[n_4-j]{\updatePcPerm{reg_3(r_1)}} \in \observations(W_5)
+We have $W_5 = \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset \uplus \ms_\act),\iota^\pwl(\dom(\ms_\unused')),\iota^\sta(\temp,\ms_f)] \futurestr W_3$, for some $\ms_\act$ and $\ms_\unused'$.
+By the execute/enter condition, we have
+                      \[
+                        \npair[n_4-j]{\updatePcPerm{reg_3(r_1)}} \in \stder(W_5)
                       \]
@@
                       \item $n_5 \leq n_4-j-2$
-                      \item $W_6 \futurewk W_5$ (as defined in the Hyp-Callee case).
-                      \item $\memSat[n_4]{\ms_6}{\revokeTemp{W_6}}$
+                      \item $W_6 \futurewk \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset \uplus \ms_\act),\iota^\pwl(\dom(\ms_\unused'')),\iota^\sta(\temp,\ms_f)]$ 
+                      \item $\memSat[n_5]{\ms_6}{\revokeTemp{W_6}}$
                       \item $\reg_6(\pcreg) = c_{\var{next}}$, $\reg_6(r_0) = \reg_3(r_0)$, $\reg_6(r_1) = \reg_3(r_1)$, $\reg(r_\env) = c_\env$
-                      \item \pointstostack{\reg_6}{\emptyset}{\ms_\unused'}
-                      \end{itemize}
+                      \item \pointstostack{\reg_6}{\emptyset}{\ms_\unused''}
+                      \end{itemize}
+                      Show 
+                      \[
+                        \npair[n_5]{(\reg_6,\ms_6 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused'')} \in \observations(W_6)
+                      \]
                       Use the $\observations$-anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) followed by the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
@@
                       \[
-                        (\reg_6,\ms_6 \uplus \ms_\unused''' \uplus \ms_f''') \step[k] (\reg_7,\ms_6\update{x}{1} \uplus \ms_\unused'' \uplus \ms_f''')
+                        (\reg_6,\ms_6 \uplus \ms_r \uplus \ms_\unused'' \uplus \ms_f''') \step[k] (\reg_7,\ms_6\update{x}{1} \uplus \ms_r \uplus \ms_\unused'' \uplus \ms_f''')
                       \]
@@
                       \[
-                        \memSat[n_4-k]{\ms_6\update{x}{1}}{W_7}
+                        \memSat[n_5-k]{\ms_6\update{x}{1}}{W_7}
                       \]
@@
         \[
-          \npair[n_4-k]{(\reg_7,\ms_6\update{x}{1} \uplus \ms_\unused'')} \in \observations(W_7)
+          \npair[n_5-k]{(\reg_7,\ms_6\update{x}{1} \uplus \ms_r \uplus \ms_\unused''')} \in \observations(W_7)
         \]
@@
         \begin{enumproof}
-          \item $\memSat[n_4-k]{\ms_6\update{x}{1}}{\revokeTemp{W_7}}$, follows from $W_7= \revokeTemp{W_7}$.
+          \item $\memSat[n_5-k]{\ms_6\update{x}{1}}{\revokeTemp{W_7}}$, follows from $W_7= \revokeTemp{W_7}$.
           \item Hyp-Callee
@@
             % TODO
-          \item Hyp-Cont
+          \item Hyp-Cont\\
+            Assume:
+            \begin{itemize}
+            \item $n_6 \leq n_5-k-2$
+            \item $W_8 \futurewk \revokeTemp{W_7}[\iota^{\sta}(\temp,\ms_\act)][\iota^\sta(\temp,\ms_{\mathit{unused}}^{(5)})][\iota^\sta(\temp,\ms_r)]$
+            \item $\memSat[n_6]{\ms_7}{\revokeTemp{W_8}}$ 
+            \item $\reg_8(r_0,r_\env) = \reg_7(r_0,r_\env)$
+            \item $\reg_8(\pcreg) = c_{\var{next}}'$
+            \item $\reg_8 \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}^{(6)}$ unused}$ for some $\ms_{\mathit{unused}}^{(6)}$
+            \end{itemize}
+            Show:
+            \[
+              \npair[n_6]{(\reg_8,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused^{(5)})} \in \observations(W_8)
+            \]
+            Use Lemma~\ref{lem:anti-red-obs}. Let $\ms_f$ be given, then
+            \[
+              (\reg_8,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused^{(5)}) \step[l] (reg_9,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0)
+            \]
+            where
+            \begin{itemize}
+            \item $\reg_9(\pcreg) = \updatePcPerm{\reg_3(r_0)}$ (note $\reg_8(r_0) = \reg_3(r_0)$)
+            \item $\reg_9(r_0) = reg_3(r_0)$
+            \item For all $r \not\in \{\pcreg,r_0\}$, $\reg_9(r) = 0$.
+            \item $\dom(\ms_0) = \dom(\ms_\unused^{(5)})$ and $\forall a \in \dom(\ms_0) \ldotp \ms_0(a) = 0$
+            \end{itemize}
+            The execution proceeds as above because $\iota_x$ in $W_8$ is in state $1$, so $\ms_7(x) = 1$ which causes the assertion to succeed.
+Subsequently the stack and most of the registers are cleared.
+
+            Now take $W_{10}$ to be $W_9$ with all the regions in $\dom(\erase{W_3}{\temp})$ reinstated. Now we show the following:
+            \begin{enumproof}
+            \item $W_{10} \futurewk W_3$\\
+              We have
+              \[
+                \forall r \in \dom(W_3) \ldotp W_3 = W_{10}
+              \]
+              if the region is permanent, then it is there because $W_{10} \futurestr W_3$. If it is temporary, then it is there because it was just reinstated.
+
+              All the future worlds we have been given have been public, so the regions can only have made public transitions.
+In $W_3$ region $\iota_x$ is in state 0 or 1.
+In $W_{10}$ region $\iota_x$ is in state 1.
+State 1 can be reached from 0 and 1 using a public transition, so the $\iota_x$ in $W_{10}$ is a public future region of the $\iota_x$ in $W_3$.
+
+              In other words, all the regions in $W_3$ have only taken public transitions compared to the corresponding regions in $W_{10}$.
+
+            \item $\memSat[n_6-l]{\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0}{W_{10}}$
+            \item $\npair[n_6-l]{\reg_9} \in \stdrr(W_{10})$
+            \end{enumproof}
             % From c_stk in val rel of W_3 get that W_3 must have iota pwl region.
@@
 
-Define the function $\activeReg{\cdot}$ as follows
+Define the function $\active{}$ as follows
 \begin{align*}
@@
 \lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
-\dominique{The comment immediately above is outdated, right?}
 Revoking all temporary regions:
@@
   \item $\memSat[n]{\ms}{\revokeTemp{W}}$ 
-  \item $\dom(\ms_f) \cup (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
+  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
   \item $(\reg,\ms) \text{ is looking at }
@@
     \item $\memSat[n-1]{\ms''}{W'}$
-    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
+    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
     \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
@@
     \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unused}}'')][\iota^\sta(\temp,\ms_f)]$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus \ms_\act)][\iota^\sta(\temp,\ms_{\mathit{unused}}'')][\iota^\sta(\temp,\ms_f)]$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
   \end{align*}
-  \dominique{The above still needs updating now that revoked regions remember
-    what they used to look like.}
   For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$, we have $W' \futurestr W$. Now define the following memory segment:

### 2017-01-25 local-capability-machine.tex
Finished main part of awkward example proof (some details missing, see todos). Changed the revoked regions so they can become anything in private future world. (forgot to commit)

@@
             \]
-            Use Lemma~\ref{lem:anti-red-obs}. Let $\ms_f$ be given, then
-            \[
-              (\reg_8,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused^{(5)}) \step[l] (reg_9,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0)
+            Use Lemma~\ref{lem:anti-red-obs}. Let $\ms_f^{(4)}$ be given, then
+            \[
+              (\reg_8,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused^{(5)} \uplus \ms_f^{(4)}) \step[l] (reg_9,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0 \uplus \ms_f^{(4)})
             \]
@@
             \begin{enumproof}
-            \item $W_{10} \futurewk W_3$\\
+            \item $W_{10} \futurewk W_3$ \label{g1:w10futwkw3}\\
               We have
@@
               \]
-              if the region is permanent, then it is there because $W_{10} \futurestr W_3$. If it is temporary, then it is there because it was just reinstated.
+              if the region was permanent in $W_3$, then it is there because $W_{10} \futurestr W_3$.
+If it was temporary, then it is there because it was just reinstated.
+If it was revoked in $W_3$, then it is still there because the only reinstated island were the temporary ones in $W_3$.
 
@@
 
-            \item $\memSat[n_6-l]{\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0}{W_{10}}$
-            \item $\npair[n_6-l]{\reg_9} \in \stdrr(W_{10})$
-            \end{enumproof}
-            % From c_stk in val rel of W_3 get that W_3 must have iota pwl region.
-            % 
+            \item $\memSat[n_6-l]{\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0}{W_{10}}$ \label{g1:memsatw10}\\
+              First notice that from %TODO write more about why this is the case. Lemma?
+              \begin{itemize}
+              \item $\npair[n_4]{\reg_3} \in \stdrr(W_3)$
+              \item $\memSat[n_4]{\ms_3}{W_3}$
+              \item $\reg(r_\stk).\perm = \rwlx$
+              \end{itemize}
+              we get that there exists a region, $r_{\adv \stk}$ in $W_3$ such that
+              \[
+                W_3(r_{\adv \stk}) = \iota_{\stk_a,\stk_b}^\pwl
+              \]
+              and $\dom(\ms_\unused) \subseteq [\stk_a,\stk_b]$. Now take $\ms_{\adv \stk} = \ms_r|_{[\stk_a,\stk_b]}$ (notice this not all of $[\stk_a,\stk_b]$ is in the domain of $\ms_r$).
+              We know
+              \begin{equation}
+                \label{eq:g1:ms7heapsat}
+                \memSat[n_6]{\ms_7}{\revokeTemp{W_8}}
+              \end{equation}
+              and
+              \begin{equation}
+                \label{eq:g1:ms3heapsat}
+                \memSat[n_4]{\ms_3}{W_3}
+              \end{equation}
+              which gives us two partitions say $P_8$ and $P_3$ respectively. Now define the partition $P$ as follows:
+              \[
+                P(r) =
+                \begin{cases}
+                  P_8(r) & r \in \dom(\erase{W_8}{\perma}) \\
+                  \ms_{\adv \stk} \uplus \ms_0 & r = r_{\adv \stk}\\
+                  P_3(r) & \text{otherwise}
+                \end{cases}
+              \]
+              Now let $r \in \active{W}$, $n_7 < n_6 - l$, and $W(r) = (\_,s,\_,\_,H)$ and show
+              \[
+                \npair[n_7]{P(r)} \in H(s)(\xi^{-1}(W_{10})).
+              \]
+              Consider the following cases
+              \begin{enumproof}
+                \item $r \in \dom(\erase{W_8}{\perma})$ \\
+                  Use \ref{eq:g1:ms7heapsat}, the fact that $W_{10} \futurestr \revokeTemp{W_8}$ and that permanent regions respect future private world.
+                \item $r = r_{\adv \stk}$ \\
+                  In this case we know the region is $\iota^\pwl_{\stk_a,\stk_b}$, so we need to show
+                  \[
+                    \npair[n_7]{\ms_{\adv \stk} \uplus \ms_0} \in H^\pwl_{\stk_a,\stk_b}(1)(\xi^{-1}(W_{10}))
+                  \]
+                  which amounts to showing
+                  \[
+                    \forall a \in \dom(\ms_0) \ldotp  \npair[n_7-1]{\ms_0(a)} \in \stdvr(W_{10}),
+                  \]
+                  which is trivial, and
+                  \[
+                    \forall a \in \dom(\ms_{\adv \stk}) \ldotp \npair[n_7-1]{\ms_{\adv \stk}(a)} \in \stdvr(W_{10})
+                  \]
+                  here we use that \ref{eq:g1:ms3heapsat} entails
+                  \[
+                    \forall a \in \dom(\ms_{\adv \stk}) \ldotp \npair[n_4-1]{\ms_{\adv \stk}(a)} \in \stdvr(W_3)
+                  \]
+                  and the fact that $\stdvr$ is monotone w.r.t $\futurewk$, $W_{10} \futurewk W_3$, and $\stdvr(W_{10})$ is downwardsclosed.
+                \item otherwise \\
+                  Use \ref{eq:g1:ms3heapsat}, $W_{10} \futurewk W_3$, and the fact that for a temporary region $H(s)$ is monotone w.r.t. $\futurewk$.
+              \end{enumproof}
+
+            \item $\npair[n_6-l]{\reg_9} \in \stdrr(W_{10})$ \label{g1:reg9stdrr}\\
+              Most registers are cleared. The only interesting register is $r_0$, so show:
+              \[
+                \npair[n_6-l]{\reg_9(r_0)}\in \stdvr(W_{10})
+              \]
+              This follows from $\reg_9(r_0) = \reg_3(r_0)$, $\npair[n_4]{\reg_3} \in \stdrr(W_3)$, $\stdvr$ monotone w.r.t $\futurewk$, $W_{10} \futurewk W_3$.
+            \end{enumproof}
+            As we were using Lemma~\ref{lem:anti-red-obs}, we need to show
+            \[
+              \npair[n_6-l]{(reg_9,\ms_7 \uplus \ms_r \uplus \emptyset \uplus \ms_0)} \in \observations(W_{10})
+            \]
+            To this end the use $\reg_3(r_0) = \reg_9(r_0)$ and $\npair[n_4]{\reg_3(r_0)} \in \stdvr1(W_3)$.
+Assuming that $\reg_9(r_0).\perm \in \{\entry,\exec,\rwx,\rwlx \}$ (if this is not the case, then it is trivial to show the above as the execution fails), then either the $\execCond{}$ or the $\entryCond{}$ hold for approriate values.
+Now use that $n_6-l < n_4$ and $W_{10} \futurewk W_3$ (\ref{g1:w10futwkw3})\footnote{We don't know whether the capability is local or global, but it does not matter as we have a public future world relation between the two worlds.} to get
+            \[
+              \npair[n_6-l]{\updatePcPerm{\reg_9(r_0)}} \in \stder(W_{10})
+            \]
+            now using \ref{g1:memsatw10} and \ref{g1:reg9stdrr}, we get the desired result.
         \end{enumproof}
@@
 \begin{align*}
-  \xi_\pub :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
+  \xi_\pub :   \Worwk \cong \blater (\nats \finparfun (   & \{\revoked\} + \\
                                                                 & \{\temp\} \times \States \times \Rels \times (\States \fun (\Worwk \monnefun \UPred{\HeapSegments})) + \\
@@
   \begin{align*}
-    \xi : \Wor \cong \blater (\nats \finparfun ( & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
+    \xi : \Wor \cong \blater (\nats \finparfun ( & \{\revoked\}  + \\
                                           & \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) + \\
@@
 \begin{align*}
-  \Regions = & \{\revoked\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
+  \Regions = & \{\revoked\} \uplus \\
              & \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
@@
   \inferrule{ }
-  { (\revoked,s,\phi_\pub,\phi,H) \futurestr (\revoked,s,\phi_\pub,\phi,H) }
+  { \revoked \futurestr (\temp,s,\phi_\pub,\phi,H) }
   \and
-  \inferrule{ (s,s') \in \phi }
-  { (\revoked,s',\phi_\pub,\phi,H) \futurestr (\temp,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ (s,s') \in \phi }
-  { (\temp,s',\phi_\pub,\phi,H) \futurestr (\revoked,s,\phi_\pub,\phi,H) }
+  \inferrule{ r \in \Regions }
+  { r \futurestr \revoked }
 \end{mathpar}
@@
   \inferrule{ }
-  { (\revoked,s,\phi_\pub,\phi,H) \futurewk (\revoked,s,\phi_\pub,\phi,H) }
+  { \revoked \futurewk \revoked }
 \end{mathpar}
@@
                    \begin{cases}
-                     (\revoked,s,\phi_\pub,\phi,H)            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
                      W(r)                & \text{otherwise}

### 2017-01-25 local-capability-machine.tex

@@
   \begin{align*}
-    & \forall n \ldotp \forall \ms, \ms' \ldotp \forall W, W' \ldotp\\
-    & \quad   \memSat{\ms}{W} \land \memSat{\ms'}{W'} \Rightarrow \\
+    & \forall n \ldotp \forall \ms, \ms', \ms'' \ldotp \forall W, W', W'' \ldotp\\
+    & \quad \ms'' = \ms \uplus \ms' \land W'' = W \uplus W' \land \memSat{\ms}{W} \land \memSat{\ms'}{W'} \Rightarrow \\
     & \qquad \heapSat[\ms \uplus \ms' ]{n}{W \uplus W'}

### 2017-01-25 local-capability-machine.tex
missing n in scall lemma
Added proof of anti reduction lemma.

@@
   \begin{itemize}
-    \item $(\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}) \in \observations(W)$
+    \item $\npair{(\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}})} \in \observations(W)$
   \end{itemize}
@@
   \begin{align*}
-    & \forall n,i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
-    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \npair[n-i]{(\reg',\ms')} \in \observations(W') \land W' \futurestr W \\
+    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
+    & \quad n' \geq n - i \land W' \futurestr W \land \\
+    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
+    & \quad \npair[n-i]{(\reg',\ms')} \in \observations(W') \\
     & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:anti-red-obs}]
+  Assume
+  \begin{enumproof} 
+    \item $n' \geq n - i$ \label{ass:anti-red-steps}
+    \item $W_2 \futurestr W_1$ \label{ass:anti-red-ftrworld}
+    \item $(\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f))$ \label{ass:anti-red-exec}
+    \item $\npair[n']{(\reg',\ms')} \in \observations(W_2)$ \label{ass:anti-red-obs}
+  \end{enumproof}
+  Show 
+  \[
+    \npair{(\reg,\ms \uplus \ms_r)}  \in \observations(W_1)
+  \]
+  To this end let $\ms_{\var{frame}}$, $m'$ and $j$ be given and assume
+  \begin{equation}
+    \label{ass:anti-red-termination}
+    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[j] (\halted,m')
+  \end{equation}
+  From \ref{ass:anti-red-exec} instantiated with $\ms_{\var{frame}}$ we know
+  \begin{equation}
+    \label{ass:anti-red-part-exec}
+    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[i] 
+  \end{equation}
+  Using \ref{ass:anti-red-termination} and \ref{ass:anti-red-part-exec}, we get 
+  \[
+    (\reg',\ms' \uplus \ms_r \uplus \ms_{\var{frame}}) \step[j-i] (\halted,m')
+  \]
+  Using this with \ref{ass:anti-red-obs} and $\ms_r \uplus \ms_{\var{frame}}$ as frame, we get $W_3 \futurestr W_2$, $\ms''$ and $\ms_{\var{rev}}$ such that
+  \begin{enumproof}[resume]
+    \item $m' = \ms'' \uplus \ms_{\var{rev}} \uplus (\ms_r \uplus \ms_{\var{frame}})$ \label{ass:anti-red-mem-split}
+    \item $\memSat[n'-(j-i)]{\ms''}{W_3}$ \label{ass:anti-red-mem-sat}
+  \end{enumproof}
+  Now use $\ms_r \uplus \ms_{\var{rev}}$ as the ``revoked'' memory, $\ms''$ as the memory that satisfies some invariants, and $W_3$ as the desired world, then \ref{ass:anti-red-mem-split} gives us the split and by downwards closure \ref{ass:anti-red-mem-sat} gives us the desired memory satisfaction.
+\end{proof}
 

### 2017-01-30 local-capability-machine.tex
Big commit - added a couple of lemmas and did 'half' of the scall lemma.

@@
     \item $W' =
-      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}')),\iota^\sta (\temp,\ms_f)]$,
+      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
     \item $\memSat[n-1]{\ms''}{W'}$
     \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
-    \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
     \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
@@
       \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
-    \item $(n,c_{\mathit{ret}}) \in \stdvr(W')$
+    \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
+    \item $\npair[n-1]{c_\stk} \in \stdvr(W')$
     \end{itemize}
@@
     \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus \ms_\act)][\iota^\sta(\temp,\ms_{\mathit{unused}}'')][\iota^\sta(\temp,\ms_f)]$
+    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus \ms_\act \uplus \ms_f),\iota^\sta(\temp,\ms_{\mathit{unused}}'')]$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
 \begin{proof}
-  Apply Hyp-Callee with $\ms_{\mathit{act}} = ...$. 
+  Assume $n$ is sufficiently large to execute all the steps up to and including the jump of $\scall{r}{\overline{r_{\var{arg}}}}{\overline{r_{\var{priv}}}}$.
+If this is not the case, then given any frame the execution will not halt successfully fast enough.
+
+  Further assume
+  \begin{enumproof}
+  \item $\memSat[n]{\ms}{\revokeTemp{W}}$ \label{pf:scall-works:mem-sat-ms}
+  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
+  \item $(\reg,\ms) \text{ is looking at }
+    \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
+    \overline{r_{\mathit{priv}}})) \text{ followed by } c_{\mathit{next}}$
+  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
+  \item Hyp-Callee \label{pf:scall-works:item:Hyp-Callee}
+  \item Hyp-Cont\label{pf:scall-works:item:Hyp-Cont}
+  \end{enumproof}
+  Now we wish to apply Lemma~\ref{lem:anti-red-obs}. To this end let $\ms_{\var{frame}}$ be given. Executing the \texttt{scall} gives us
+  \[
+    (\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}} \uplus \ms_f) \step[i] (\reg_1,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused' \uplus \ms_{\var{frame}})
+  \]
+  where
+  \begin{enumproof}[resume]
+  \item $i \leq n$
+  \item $\ms_\act$ contains activation record, $\reg(\overline{r_{\var{priv}}})$, the code return capability, and the full stack capability ($\reg(r_\stk)$ with the pointer adjusted).
+  \item $\forall a \in \dom(\ms_\unused') \ldotp \ms_\unused(a) = 0$
+  \item $\dom(\ms_\unused) = \dom(\ms_\act \uplus \ms_\unused)$
+  \item $\reg'(r_0) = c_{\var{ret}} = ((\entry,\local),\_,\_,\_)$ where the range of authority is the same as $\reg(r_\stk)$ and it points to the first instruction of the activation code.
+  \item \pointstostack{\reg'}{\emptyset}{\ms_\unused'}
+  \item $\reg'(\pcreg) = \updatePcPerm{\reg(\pcreg)}$ 
+  \item $\reg'(r) = \reg(r)$
+  \item $\reg'(\overline{r_{\var{args}}}) = \reg(\overline{r_{\var{args}}})$
+  \item $\forall r' \in \RegName \setminus \{\pcreg, r_\stk, r, \overline{r_{\var{args}}} \}$
+  \end{enumproof}
+  In order to use Lemma~\ref{lem:anti-red-obs}, we now need to show
+  \[
+    \npair[n_1]{(\reg_1,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused')} \in \observations{W_1}
+  \]
+  where
+  \[
+    W_1 = \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]
+  \]
+  to this end use Hyp-Callee (\ref{pf:scall-works:Hyp-Callee}). To use this everything is satisfied directly by assumptions but the following:
+  \begin{enumproof}[resume]
+    \item $\memSat[n-1]{\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused'}{W_1}$\\
+      Here we apply Lemma~\ref{lem:disj-mem-sat}. By assumption~\ref{pf:scall-works:mem-sat-ms} we have $\memSat[n]{\ms}{\revokeTemp{W}}$. So it suffices to show
+      \[
+        \memSat[n-1]{\ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused'}{[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]}
+      \]
+      This turns out to be trivial as $\ms_f$, $\ms_\stk$, and $\ms_{\var{act}}$ match the static region. $\ms_\unused'$ is all zeroes, to it trivially satisfies the $\iota^\pwl$ region.
+    \item $\npair[n-1]{\reg'(r_\stk)} \in \stdvr(W_1)$ \\
+      Say $\dom(\reg'(r_\stk))=[\start_\stk,\addrend_\stk$. This amounts to showing
+      \begin{enumproof}
+      \item $\npair[n-1]{(\start_\stk, \addrend_\stk)} \in \readCond{}(\local)(W)$ \label{pf:scall-works:readcond-stack}\\
+        This amounts to showing 
+        \[
+          \iota^\pwl_{\start_\stk, \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk, \addrend_\stk}
+        \]
+        which is indeed the case.
+      \item $\npair[n-1]{(\start_\stk, \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$ \label{pf:scall-works:writecond-stack}\\
+        This amounts to showing 
+        \[
+          \iota^\pwl_{\start_\stk, \addrend_\stk} \nsupsim[n-1] \iota^\pwl_{\start_\stk, \addrend_\stk}
+        \]
+        which is indeed the case.
+      \item $\npair[n-1]{(\rwlx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+        Using \ref{pf:scall-works:writecond-stack} and \ref{pf:scall-works:readcond-stack}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
+      \item $\npair[n-1]{(\rwx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+        Using \ref{pf:scall-works:writecond-stack} and \ref{pf:scall-works:readcond-stack}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
+      \item $\npair[n-1]{(\exec,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+        Using \ref{pf:scall-works:writecond-stack} and \ref{pf:scall-works:readcond-stack}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
+      \end{enumproof}
+    \item $\npair[n-1]{c_{\var{ret}}} \in \stdvr(W_1)$
+      % TODO
+  \end{enumproof}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:stack-exec-cond-help}
+  If
+  \begin{itemize}
+  \item $\perm \in \{\exec,\rwx,\rwlx \}$
+  \item $\npair{(\start,\addrend)} \readCond{}(\local)(W)$
+  \item $\npair{(\start,\addrend)} \writeCond{}(\iota^\pwl,\local)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair[n]{\perm,\start,\addrend} \in \execCond{}(\local)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stack-exec-cond-help}]
+Assume
+\begin{enumproof}
+\item $\perm \in \{\exec,\rwx,\rwlx \}$
+\item $\npair{(\start,\addrend)} \readCond{}(\local)(W)$ \label{pf:stack-exec:ass:readcond}
+\item $\npair{(\start,\addrend)} \writeCond{}(\iota^\pwl,\local)(W)$ \label{pf:stack-exec:ass:writecond}
+\end{enumproof}
+Let $W' \futurewk W$, $\addr$, and $n' \leq n$ be given and show
+\[
+  \npair[n']{((\perm,\local),\start,\addrend,\addr)} \in \stder(W')
+\]
+Consider each of the three cases for $\perm$:
+\begin{enumproof}[resume]
+  \item $\perm = \rwlx$\\
+    In this case $\iota= \iota^\pwl$. If we use the FTLR (Theorem~\ref{thm:ftlr}), then we are done. It suffices to show:
+    \begin{enumproof}
+      \item $\npair[n']{(\start,\addrend) \readCond{}(\local)(W')}$ \\
+        Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:readcond}.
+      \item $\npair[n']{(\start,\addrend) \writeCond{}(\iota^\pwl,\local)(W')}$\\
+        Follows from Lemma~\ref{lem:writecond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:writecond}.
+    \end{enumproof}
+  \item $\perm = \exec$\\
+    In this case $\iota= \iota^\nwl$. If we use the FTLR (Theorem~\ref{thm:ftlr}), then we are done. It suffices to show:
+    \begin{enumproof}
+      \item $\npair[n']{(\start,\addrend) \readCond{}(\local)(W')}$ \\
+        Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:readcond}.
+      \item $\npair[n']{(\start,\addrend) \writeCond{}(\iota^\nwl,\local)(W')}$\\
+        Follows from Lemma~\ref{lem:wc-pwl-implies-wc-nwl}, Lemma~\ref{lem:writecond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:writecond}.
+    \end{enumproof}
+  \item $\perm = \rwx$\\
+    In this case $\iota= \iota^\nwl$. If we use the FTLR (Theorem~\ref{thm:ftlr}), then we are done. It suffices to show:
+    \begin{enumproof}
+      \item $\npair[n']{(\start,\addrend) \readCond{}(\local)(W')}$ \\
+        Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:readcond}.
+      \end{enumproof}
+\end{enumproof}
 \end{proof}
@@
 \lau{$a \leq b$ that the interpretation is not empty.}
+
+\subsubsection*{Permission based conditions}
+\begin{lemma}
+  \label{lem:wc-pwl-implies-wc-nwl}
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W) \Rightarrow \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\local)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of lemma~\ref{lem:wc-pwl-implies-wc-nwl}]
+  Follows from Lemma~\ref{lem:nwlp-subset-pwl}.
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:readcond-mono-pub}
+  \[
+    \forall W' \futurewk W \ldotp \npair{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair{\start,\addrend} \in \readCond{}(\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
+  \label{lem:readcond-dc}
+  \[
+    \forall n' \leq n \ldotp \npair[n]{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair[n']{\start,\addrend} \in \readCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:writecond-mono-pub}
+  \[
+    \forall W' \futurewk W \ldotp \forall \iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\} \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W) \Rightarrow \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
+  \label{lem:writecond-mono-pub}
+  \[
+    \forall n' \leq \ldotp \forall \iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\} \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W) \Rightarrow \npair[n']{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
+  % TODO
+\end{proof}
 

### 2017-01-31 local-capability-machine.tex
addressed a couple of comments

@@
       \item case $\npair[n']{\ms_\stk'|_{[\start_\stk, \start_\stk''-1]}} \in
-        H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})(W)$ \dominique{apply to world?}:\\ This region matches only this memory segment.
+        H^\sta(\ms_\stk'|_{[\start_\stk,\start_\stk''-1]})(1)(W)$ :\\ This region matches only this memory segment.
       \item case $\npair[n']{\ms_\stk'|_{[\start_\stk'', \addrend_\stk]}} \in H_{\start_\stk'', \addrend_\stk}(1)(W)$:\\ Show $\forall a \in [\start_\stk'', \addrend_\stk] \ldotp \ms_\stk'|_{[\start_\stk'', \addrend_\stk]}(a) \in \stdvr(W)$, trivially true since these memory cells contain 0.
@@
 \end{align*}
-\lau{What is the reason to have the ``view'' always be $\temp$? It seems like it does not play well with the read condition if it is always $\temp$.}
-\dominique{The comment immediately above is outdated, right?}
 Revoking all temporary regions:

### 2017-01-31 local-capability-machine.tex
Add some captions, move the antireduction lemma forward.
Start work on FTLR, move some more lemmas forward in the document, fix a problem in the definition of the observation predicate.

@@
 
-\section{Domains and Notation}
+\section{Capability Machine Definition and Operational Semantics}
+\subsection{Domains and Notation}
 
@@
 
-\subsection*{Semantics}
+\subsection{Operational Semantics}
 Assume a $\decode$ function that decodes words to instructions:
@@
 \]
+
+We define the operational semantics as follows:
 \begin{align*}
@@
                                & \forall \ms_f, \heap', i \leq n \ldotp \\
-                               & \quad(\reg,\hs \uplus \ms_f) \step[i] (\halted,\heap')  \\
+                               & \quad(\reg,\hs \uplus \ms_f) \step[i] (\halted,\Phi)  \\
                                & \qquad \Rightarrow
@@
                                  & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
-                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \ms_f \land \\ 
+                                 & \quad \memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f \land \\ 
                                  & \quad \heapSat[\hs']{n-i}{W'} \}
@@
 
-\begin{theorem}[Fundamental theorem of logical relations]
-\lau{do we want the FTLR to also talk about everything that is trivially in the FTLR.}
-  \label{thm:ftlr}
+\subsubsection{Antireduction for the observation relation}
+
+\begin{lemma}[Anti-reduction for $\observations$]
+\label{lem:anti-red-obs}
+  \begin{align*}
+    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
+    & \quad n' \geq n - i \land W' \futurestr W \land \\
+    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
+    & \quad \npair[n-i]{(\reg',\ms')} \in \observations(W') \\
+    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:anti-red-obs}]
+  Assume
+  \begin{enumproof} 
+    \item $n' \geq n - i$ \label{ass:anti-red-steps}
+    \item $W_2 \futurestr W_1$ \label{ass:anti-red-ftrworld}
+    \item $(\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f))$ \label{ass:anti-red-exec}
+    \item $\npair[n']{(\reg',\ms')} \in \observations(W_2)$ \label{ass:anti-red-obs}
+  \end{enumproof}
+  Show 
+  \[
+    \npair{(\reg,\ms \uplus \ms_r)}  \in \observations(W_1)
+  \]
+  To this end let $\ms_{\var{frame}}$, $m'$ and $j$ be given and assume
+  \begin{equation}
+    \label{ass:anti-red-termination}
+    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[j] (\halted,m')
+  \end{equation}
+  From \ref{ass:anti-red-exec} instantiated with $\ms_{\var{frame}}$ we know
+  \begin{equation}
+    \label{ass:anti-red-part-exec}
+    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[i] 
+  \end{equation}
+  Using \ref{ass:anti-red-termination} and \ref{ass:anti-red-part-exec}, we get 
+  \[
+    (\reg',\ms' \uplus \ms_r \uplus \ms_{\var{frame}}) \step[j-i] (\halted,m')
+  \]
+  Using this with \ref{ass:anti-red-obs} and $\ms_r \uplus \ms_{\var{frame}}$ as frame, we get $W_3 \futurestr W_2$, $\ms''$ and $\ms_{\var{rev}}$ such that
+  \begin{enumproof}[resume]
+    \item $m' = \ms'' \uplus \ms_{\var{rev}} \uplus (\ms_r \uplus \ms_{\var{frame}})$ \label{ass:anti-red-mem-split}
+    \item $\memSat[n'-(j-i)]{\ms''}{W_3}$ \label{ass:anti-red-mem-sat}
+  \end{enumproof}
+  Now use $\ms_r \uplus \ms_{\var{rev}}$ as the ``revoked'' memory, $\ms''$ as the memory that satisfies some invariants, and $W_3$ as the desired world, then \ref{ass:anti-red-mem-split} gives us the split and by downwards closure \ref{ass:anti-red-mem-sat} gives us the desired memory satisfaction.
+\end{proof}
+
+\subsubsection{Standard regions}
+\begin{lemma}
+  \label{lem:nwl-subset-pwl}
+  \begin{align*}
+    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
+    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
+  \end{align*}
+\end{lemma}
+\begin{lemma}
+  \label{lem:nwlp-subset-pwl}
+  \begin{align*}
+    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
+    & \quad \iota^{\nwl,p}_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
+  \end{align*}
+\end{lemma}
+
+\begin{lemma}
+  \label{lem:pwl-nsim-view}
+  \begin{align*}
+    & \forall n \in \nats \ldotp \forall \start, \addrend, b \in \Addrs \ldotp \forall \iota \in \Regions \\
+    & \iota \nsim \iota^\pwl_{\start,\addrend} \land \start \leq \addrend \Rightarrow \iota \nequal \iota^\pwl_{\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:pwl-nsim-view}]
+  For $n = 0$ it is trivial, so assume $n > 0$.
+  Say $\iota = (v,s,\phi_{\var{pub}},\phi,H)$, then by $\nsim$, we know $s = 1$, $\phi_{\var{pub}} \equiv \phi \equiv =$, and $H = H^\pwl_{\start,\addrend}$.
+It remains to show that $v=\temp$.
+To do so, we show that it cannot be the case that $v=\perma$.
+If $v = \perma$, then $H$ must be monotone with respect to $\futurestr$.
+If we can show that this is not the case, then for $\iota$ to be a region it must be the case that $v \neq \perma$ and thus $v = \temp$.
+
+  To this end let $b\not\in[\start,\addrend]$ and define the worlds:
+  \begin{align*}
+    \xi (W)  =& [0 \mapsto \iota^\pwl_{\start,\addrend}]  \\ 
+              & [1 \mapsto \iota^\pwl_{b,b}] \\
+    \xi (W')  =& [0 \mapsto \iota^\pwl_{\start,\addrend}]  \\ 
+              & [1 \mapsto \revoked] 
+  \end{align*}
+  For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$, we have $W' \futurestr W$. Now define the following memory segment:
+  \[
+    \ms = [\start \mapsto ((\readonly,\local),b,b,b), \start+1 \mapsto 0, \dots , \addrend \mapsto 0]
+  \]
+  It is the case that
+  \[
+    \npair{\ms} \in H \; 1 \; W
+  \]
+  but
+  \[
+    \npair{\ms} \not\in H \; 1 \; W'
+  \]
+  as it is not the case that \lau{what if $n=1$? Need to figure out when exactly this holds}
+  \[
+    \npair[n-1]{((\readonly,\local),b,b,b)} \in \stdvr(\xi(W')).
+  \]
+  The only other option that remains is $v=\temp$.
+\end{proof}
+\lau{$a \leq b$ that the interpretation is not empty.}
+
+\subsubsection{Permission based conditions}
+\begin{lemma}
+  \label{lem:wc-pwl-implies-wc-nwl}
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W) \Rightarrow \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\local)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of lemma~\ref{lem:wc-pwl-implies-wc-nwl}]
+  Follows from Lemma~\ref{lem:nwlp-subset-pwl}.
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:readcond-mono-pub}
+  \[
+    \forall W' \futurewk W \ldotp \npair{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair{\start,\addrend} \in \readCond{}(\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
+  \label{lem:readcond-dc}
+  \[
+    \forall n' \leq n \ldotp \npair[n]{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair[n']{\start,\addrend} \in \readCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:writecond-mono-pub}
+  \[
+    \forall W' \futurewk W \ldotp \forall \iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\} \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W) \Rightarrow \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
+  \label{lem:writecond-dc}
+  For all $n' \leq n$, $\iota \in
+  \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$,
+  If $\npair{\start,\addrend} \in
+    \writeCond{}(\iota,\gl)(W)$
+  then 
+    $\npair[n']{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)$
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
+  % TODO
+\end{proof}
+
+\subsubsection{Fundamental theorem of logical relations}
+
+\begin{lemma}[Fundamental theorem proof helper lemma]
+  \label{lem:ftlr-induction-helper}
   For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
-  if
-  \[
+  If one of the following holds:
+  \begin{itemize}
+  \item \[
       \begin{gathered}
@@
     \]
-  or
-  \[
+  \item \[
       \begin{gathered}
@@
     \]
+  \end{itemize}
+  and
+  \begin{itemize}
+  \item $\npair[n]{\reg}\in \stdrr(W)$
+  \item $\heapSat[\hs]{n}{W}$
+  \item $i \leq n$
+  \item $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$
+  \end{itemize}
+
+  Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
+  $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
+\end{lemma}
+\begin{proof}
+  \begin{enumproof}
+  \item By induction on $n$. In other words, assume that the theorem already
+    holds for all $n' < n$.
+  \item The case that $i = 0$ is impossible, since
+    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. Assume w.l.o.g. that $i = 1+i'$
+  \item 
+  \end{enumproof}
+\end{proof}
+
+\begin{theorem}[Fundamental theorem of logical relations]
+  \label{thm:ftlr}
+  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
+  If one of the following holds:
+  \begin{itemize}
+  \item \[
+      \begin{gathered}
+        \perm = \exec \land\\
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+      \end{gathered}
+    \]
+  \item \[
+      \begin{gathered}
+        \perm = \rwx \land \\
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
+        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
+      \end{gathered}
+    \]
+  \item \[
+      \begin{gathered}
+        \perm = \rwlx \land\\
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
+        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
+      \end{gathered}
+    \]
+  \end{itemize}
   then
@@
 \end{theorem}
-
+\begin{proof}
+  \begin{enumproof}
+  \item Assume: $n' \leq n$, $\npair[n']{\reg}\in \stdrr(W)$, $\heapSat[\hs]{n'}{W}$.\\
+    Suffices: $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs)}\in\observations(W)$.\\
+    By: definition of $\stder(W)$.
+  \item Assume: $\ms_f$, $\heap'$, $i \leq n'$, $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$,\\
+    Suffices: $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
+    $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+    By: definition of $\observations(W)$ \label{suff-after-eval}
+  \item This follows by Lemma~\ref{lem:ftlr-induction-helper} using
+    Lemmas~\ref{lem:readcond-dc} and~\ref{lem:writecond-dc}.
+  \end{enumproof}
+\end{proof}
+
+\subsubsection{Scall macro-instruction correctness}
 \begin{definition}
@@
 
+\subsubsection{Helper lemmas about the stack}
 \begin{lemma}
@@
 \end{enumproof}
-\end{proof}
-
-\begin{lemma}[Anti-reduction for $\observations$]
-\label{lem:anti-red-obs}
-  \begin{align*}
-    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
-    & \quad n' \geq n - i \land W' \futurestr W \land \\
-    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
-    & \quad \npair[n-i]{(\reg',\ms')} \in \observations(W') \\
-    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
-  \end{align*}
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:anti-red-obs}]
-  Assume
-  \begin{enumproof} 
-    \item $n' \geq n - i$ \label{ass:anti-red-steps}
-    \item $W_2 \futurestr W_1$ \label{ass:anti-red-ftrworld}
-    \item $(\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f))$ \label{ass:anti-red-exec}
-    \item $\npair[n']{(\reg',\ms')} \in \observations(W_2)$ \label{ass:anti-red-obs}
-  \end{enumproof}
-  Show 
-  \[
-    \npair{(\reg,\ms \uplus \ms_r)}  \in \observations(W_1)
-  \]
-  To this end let $\ms_{\var{frame}}$, $m'$ and $j$ be given and assume
-  \begin{equation}
-    \label{ass:anti-red-termination}
-    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[j] (\halted,m')
-  \end{equation}
-  From \ref{ass:anti-red-exec} instantiated with $\ms_{\var{frame}}$ we know
-  \begin{equation}
-    \label{ass:anti-red-part-exec}
-    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[i] 
-  \end{equation}
-  Using \ref{ass:anti-red-termination} and \ref{ass:anti-red-part-exec}, we get 
-  \[
-    (\reg',\ms' \uplus \ms_r \uplus \ms_{\var{frame}}) \step[j-i] (\halted,m')
-  \]
-  Using this with \ref{ass:anti-red-obs} and $\ms_r \uplus \ms_{\var{frame}}$ as frame, we get $W_3 \futurestr W_2$, $\ms''$ and $\ms_{\var{rev}}$ such that
-  \begin{enumproof}[resume]
-    \item $m' = \ms'' \uplus \ms_{\var{rev}} \uplus (\ms_r \uplus \ms_{\var{frame}})$ \label{ass:anti-red-mem-split}
-    \item $\memSat[n'-(j-i)]{\ms''}{W_3}$ \label{ass:anti-red-mem-sat}
-  \end{enumproof}
-  Now use $\ms_r \uplus \ms_{\var{rev}}$ as the ``revoked'' memory, $\ms''$ as the memory that satisfies some invariants, and $W_3$ as the desired world, then \ref{ass:anti-red-mem-split} gives us the split and by downwards closure \ref{ass:anti-red-mem-sat} gives us the desired memory satisfaction.
 \end{proof}
@@
 
+\subsubsection{Memory Segment Satisfaction}
 We expect the following lemmas to hold true:
-\subsubsection{Memory Segment Satisfaction}
 \begin{lemma}[Downwards closure]
@@
 
-\begin{lemma}[Public monotonicity]
-  \lau{This is not true! If the new region is active and does not accept the empty memory, then it is no longer possible to make a satisfactory partition of the heap.}
-  \begin{align*}
-    & \forall \hs, n, W' \futurewk W \ldotp \\
-    & \quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n}{W'}
-  \end{align*}
-\end{lemma}
+%   \lau{This is not true! If the new region is active and does not accept the empty memory, then it is no longer possible to make a satisfactory partition of the heap.}
+%   \dominique{I agree. World satisfaction for heaps is not monotonous and it
+%   shouldn't be. Commenting out the lemma!}
+% \begin{lemma}[Public monotonicity]
+%   \begin{align*}
+%     & \forall \hs, n, W' \futurewk W \ldotp \\
+%     & \quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n}{W'}
+%   \end{align*}
+% \end{lemma}
 
@@
 
-\subsubsection{Future world}
+\subsubsection{Future worlds}
 \begin{lemma}
@@
 \end{lemma}
-
-\subsubsection{Standard regions}
-\begin{lemma}
-  \label{lem:nwl-subset-pwl}
-  \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
-  \end{align*}
-\end{lemma}
-\begin{lemma}
-  \label{lem:nwlp-subset-pwl}
-  \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \iota^{\nwl,p}_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
-  \end{align*}
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:pwl-nsim-view}
-  \begin{align*}
-    & \forall n \in \nats \ldotp \forall \start, \addrend, b \in \Addrs \ldotp \forall \iota \in \Regions \\
-    & \iota \nsim \iota^\pwl_{\start,\addrend} \land \start \leq \addrend \Rightarrow \iota \nequal \iota^\pwl_{\start,\addrend}
-  \end{align*}
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:pwl-nsim-view}]
-  For $n = 0$ it is trivial, so assume $n > 0$.
-  Say $\iota = (v,s,\phi_{\var{pub}},\phi,H)$, then by $\nsim$, we know $s = 1$, $\phi_{\var{pub}} \equiv \phi \equiv =$, and $H = H^\pwl_{\start,\addrend}$.
-It remains to show that $v=\temp$.
-To do so, we show that it cannot be the case that $v=\perma$.
-If $v = \perma$, then $H$ must be monotone with respect to $\futurestr$.
-If we can show that this is not the case, then for $\iota$ to be a region it must be the case that $v \neq \perma$ and thus $v = \temp$.
-
-  To this end let $b\not\in[\start,\addrend]$ and define the worlds:
-  \begin{align*}
-    \xi (W)  =& [0 \mapsto \iota^\pwl_{\start,\addrend}]  \\ 
-              & [1 \mapsto \iota^\pwl_{b,b}] \\
-    \xi (W')  =& [0 \mapsto \iota^\pwl_{\start,\addrend}]  \\ 
-              & [1 \mapsto \revoked] 
-  \end{align*}
-  For these two worlds, we have $\xi(W') \futurestr \xi(W)$ and from mono.\ of $\xi^{-1}$, we have $W' \futurestr W$. Now define the following memory segment:
-  \[
-    \ms = [\start \mapsto ((\readonly,\local),b,b,b), \start+1 \mapsto 0, \dots , \addrend \mapsto 0]
-  \]
-  It is the case that
-  \[
-    \npair{\ms} \in H \; 1 \; W
-  \]
-  but
-  \[
-    \npair{\ms} \not\in H \; 1 \; W'
-  \]
-  as it is not the case that \lau{what if $n=1$? Need to figure out when exactly this holds}
-  \[
-    \npair[n-1]{((\readonly,\local),b,b,b)} \in \stdvr(\xi(W')).
-  \]
-  The only other option that remains is $v=\temp$.
-\end{proof}
-\lau{$a \leq b$ that the interpretation is not empty.}
-
-\subsubsection*{Permission based conditions}
-\begin{lemma}
-  \label{lem:wc-pwl-implies-wc-nwl}
-  \[
-    \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W) \Rightarrow \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\local)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of lemma~\ref{lem:wc-pwl-implies-wc-nwl}]
-  Follows from Lemma~\ref{lem:nwlp-subset-pwl}.
-\end{proof}
-
-\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:readcond-mono-pub}
-  \[
-    \forall W' \futurewk W \ldotp \npair{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair{\start,\addrend} \in \readCond{}(\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
-  \label{lem:readcond-dc}
-  \[
-    \forall n' \leq n \ldotp \npair[n]{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair[n']{\start,\addrend} \in \readCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:writecond-mono-pub}
-  \[
-    \forall W' \futurewk W \ldotp \forall \iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\} \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W) \Rightarrow \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
-  \label{lem:writecond-mono-pub}
-  \[
-    \forall n' \leq \ldotp \forall \iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\} \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W) \Rightarrow \npair[n']{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
-  % TODO
-\end{proof}
 

### 2017-01-31 local-capability-machine.tex
Reverted observation predicate change and changed the Conf definition instead
fixed typos in the scall and stack proof. Edited some lemmas to be easier to read and stay on page

@@
   \Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
-  \Confs &::= \ExecConfs + \{\failed \} + \halted \times \ExecConfs \\
+  \Confs &::= \ExecConfs + \{\failed \} + \halted \times \Heaps \\
   \HeapSegments &::= \Addrs \parfun \Words
@@
                                & \forall \ms_f, \heap', i \leq n \ldotp \\
-                               & \quad(\reg,\hs \uplus \ms_f) \step[i] (\halted,\Phi)  \\
+                               & \quad(\reg,\hs \uplus \ms_f) \step[i] (\halted,\heap')  \\
                                & \qquad \Rightarrow
@@
                                  & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
-                                 & \quad \memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f \land \\ 
+                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \ms_f \land \\ 
                                  & \quad \heapSat[\hs']{n-i}{W'} \}
@@
   \label{lem:wc-pwl-implies-wc-nwl}
-  \[
-    \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W) \Rightarrow \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\local)(W)
+  If
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W)
+  \]
+  then
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\local)(W)
   \]
@@
   \label{lem:readcond-mono-pub}
-  \[
-    \forall W' \futurewk W \ldotp \npair{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair{\start,\addrend} \in \readCond{}(\gl)(W')
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\npair{\start,\addrend} \in \readCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{\start,\addrend} \in \readCond{}(\gl)(W')
   \]
@@
   \label{lem:readcond-dc}
-  \[
-    \forall n' \leq n \ldotp \npair[n]{\start,\addrend} \in \readCond{}(\gl)(W) \Rightarrow \npair[n']{\start,\addrend} \in \readCond{}(\gl)(W)
+  If
+  \begin{itemize}
+  \item $ n' \leq n$
+  \item $\npair[n]{\start,\addrend} \in \readCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair[n']{\start,\addrend} \in \readCond{}(\gl)(W)
   \]
@@
   \label{lem:writecond-mono-pub}
-  \[
-    \forall W' \futurewk W \ldotp \forall \iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\} \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W) \Rightarrow \npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W')
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{\start,\addrend}\in \writeCond{}(\iota,\gl)(W')
   \]
@@
   \label{lem:writecond-dc}
-  For all $n' \leq n$, $\iota \in
-  \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$,
-  If $\npair{\start,\addrend} \in
-    \writeCond{}(\iota,\gl)(W)$
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)$
+  \end{itemize}
   then 
-    $\npair[n']{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)$
+  \[
+    \npair[n']{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)
+  \]
 \end{lemma}
@@
   \item $(\reg,\ms) \text{ is looking at }
-    \mathtt{scall}(r(\overline{r_{\mathit{arg}}},
-    \overline{r_{\mathit{priv}}})) \text{ followed by } c_{\mathit{next}}$
+    \mathtt{scall}\;r(\overline{r_{\mathit{arg}}},
+    \overline{r_{\mathit{priv}}}) \text{ followed by } c_{\mathit{next}}$
   \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
@@
 \begin{proof}
-  Assume $n$ is sufficiently large to execute all the steps up to and including the jump of $\scall{r}{\overline{r_{\var{arg}}}}{\overline{r_{\var{priv}}}}$.
-If this is not the case, then given any frame the execution will not halt successfully fast enough.
+  Assume $n$ is sufficiently large to execute all the steps up to and including the jump of $\mathtt{scall}\; r(\overline{r_{\var{arg}}},\overline{r_{\var{priv}}})$.
+If this is not the case, then in any given memory frame the execution will not halt successfully fast enough.
 
@@
   \[
-    (\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}} \uplus \ms_f) \step[i] (\reg_1,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused' \uplus \ms_{\var{frame}})
+    (\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}} \uplus \ms_{\var{frame}}) \step[i] (\reg_1,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused' \uplus \ms_{\var{frame}})
   \]
@@
   \item $\ms_\act$ contains activation record, $\reg(\overline{r_{\var{priv}}})$, the code return capability, and the full stack capability ($\reg(r_\stk)$ with the pointer adjusted).
-  \item $\forall a \in \dom(\ms_\unused') \ldotp \ms_\unused(a) = 0$
-  \item $\dom(\ms_\unused) = \dom(\ms_\act \uplus \ms_\unused)$
-  \item $\reg'(r_0) = c_{\var{ret}} = ((\entry,\local),\_,\_,\_)$ where the range of authority is the same as $\reg(r_\stk)$ and it points to the first instruction of the activation code.
-  \item \pointstostack{\reg'}{\emptyset}{\ms_\unused'}
-  \item $\reg'(\pcreg) = \updatePcPerm{\reg(\pcreg)}$ 
-  \item $\reg'(r) = \reg(r)$
-  \item $\reg'(\overline{r_{\var{args}}}) = \reg(\overline{r_{\var{args}}})$
-  \item $\forall r' \in \RegName \setminus \{\pcreg, r_\stk, r, \overline{r_{\var{args}}} \}$
+  \item $\forall a \in \dom(\ms_\unused') \ldotp \ms_\unused'(a) = 0$
+  \item $\dom(\ms_\unused) = \dom(\ms_\act \uplus \ms_\unused')$
+  \item $\reg_1(r_0) = c_{\var{ret}} = ((\entry,\local),\_,\_,\_)$ where the range of authority is the same as $\reg(r_\stk)$ and it points to the first instruction of the activation code. 
+    % TODO Illustrate this part of the heap better
+  \item \pointstostack{\reg_1}{\emptyset}{\ms_\unused'} \label{pf:scall-works:item:points-to-empty-and-unusedp}
+  \item $\reg_1(\pcreg) = \updatePcPerm{\reg(\pcreg)}$ 
+  \item $\reg_1(r) = \reg(r)$
+  \item $\reg_1(\overline{r_{\var{args}}}) = \reg(\overline{r_{\var{args}}})$
+  \item $\forall r' \in \RegName \setminus \{\pcreg, r_\stk, r, \overline{r_{\var{args}}} \} \ldotp \reg_1(r') = 0$
   \end{enumproof}
@@
   \[
-    \npair[n_1]{(\reg_1,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused')} \in \observations{W_1}
+    \npair[n_1]{(\reg_1,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{act}} \uplus \ms_\unused')} \in \observations(W_1)
   \]
@@
   \]
-  to this end use Hyp-Callee (\ref{pf:scall-works:Hyp-Callee}). To use this everything is satisfied directly by assumptions but the following:
+  to this end use Hyp-Callee (\ref{pf:scall-works:item:Hyp-Callee}). To use this everything is satisfied directly by assumptions but the following:
   \begin{enumproof}[resume]
@@
     \item $\npair[n-1]{\reg'(r_\stk)} \in \stdvr(W_1)$ \\
-      Say $\dom(\reg'(r_\stk))=[\start_\stk,\addrend_\stk$. This amounts to showing
-      \begin{enumproof}
-      \item $\npair[n-1]{(\start_\stk, \addrend_\stk)} \in \readCond{}(\local)(W)$ \label{pf:scall-works:readcond-stack}\\
-        This amounts to showing 
-        \[
-          \iota^\pwl_{\start_\stk, \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk, \addrend_\stk}
-        \]
-        which is indeed the case.
-      \item $\npair[n-1]{(\start_\stk, \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$ \label{pf:scall-works:writecond-stack}\\
-        This amounts to showing 
-        \[
-          \iota^\pwl_{\start_\stk, \addrend_\stk} \nsupsim[n-1] \iota^\pwl_{\start_\stk, \addrend_\stk}
-        \]
-        which is indeed the case.
-      \item $\npair[n-1]{(\rwlx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-        Using \ref{pf:scall-works:writecond-stack} and \ref{pf:scall-works:readcond-stack}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
-      \item $\npair[n-1]{(\rwx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-        Using \ref{pf:scall-works:writecond-stack} and \ref{pf:scall-works:readcond-stack}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
-      \item $\npair[n-1]{(\exec,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
-        Using \ref{pf:scall-works:writecond-stack} and \ref{pf:scall-works:readcond-stack}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
-      \end{enumproof}
+      Use Lemma~\ref{lem:stack-cap-vr} with \ref{pf:scall-works:item:points-to-empty-and-unusedp} and that $W_1$ has region $\iota^{\pwl}(\dom(\ms_{\mathit{unused}}')$.
     \item $\npair[n-1]{c_{\var{ret}}} \in \stdvr(W_1)$
@@
     \begin{enumproof}
-      \item $\npair[n']{(\start,\addrend) \readCond{}(\local)(W')}$ \\
+      \item $\npair[n']{(\start,\addrend)} \in \readCond{}(\local)(W')$ \\
         Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:readcond}.
-      \item $\npair[n']{(\start,\addrend) \writeCond{}(\iota^\pwl,\local)(W')}$\\
+      \item $\npair[n']{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W')$\\
         Follows from Lemma~\ref{lem:writecond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:writecond}.
@@
     \begin{enumproof}
-      \item $\npair[n']{(\start,\addrend) \readCond{}(\local)(W')}$ \\
+      \item $\npair[n']{(\start,\addrend)} \in \readCond{}(\local)(W')$ \\
         Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:readcond}.
-      \item $\npair[n']{(\start,\addrend) \writeCond{}(\iota^\nwl,\local)(W')}$\\
+      \item $\npair[n']{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\local)(W')$\\
         Follows from Lemma~\ref{lem:wc-pwl-implies-wc-nwl}, Lemma~\ref{lem:writecond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:writecond}.
@@
     \begin{enumproof}
-      \item $\npair[n']{(\start,\addrend) \readCond{}(\local)(W')}$ \\
+      \item $\npair[n']{(\start,\addrend)} \in \readCond{}(\local)(W')$ \\
         Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:readcond-dc}, and assumption~\ref{pf:stack-exec:ass:readcond}.
@@
   \label{lem:stack-cap-vr}
-  \begin{align*}
-    & \pointstostack{\reg}{\emptyset}{\ms} \land \\
-    & (\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))) \\
-    & \quad \Rightarrow \npair{\reg(r_\stk)} \in \stdvr(W)
-  \end{align*}
+  If
+  \begin{itemize}
+  \item \pointstostack{\reg}{\emptyset}{\ms}
+  \item $\exists r \ldotp W(r) = \iota^\pwl (\dom(\ms))$
+  \end{itemize}
+  then
+  \[
+    \npair{\reg(r_\stk)} \in \stdvr(W)
+  \]
 \end{lemma}
@@
       which is true as they are even equal.
-    \item $\npair{(\rwlx,\start,\addrend)} \in \execCond{}(\local)(W)$ : \label{pf:stack-cap-exec}\\
-      Let $n' < n$, $W' \futurewk W$, and $a\in \dom(\start,\addrend)$ be given and show
-      \[
-        \npair[n']{((\rwlx,\local),\start,\addrend,a)} \in \stder(W')
-      \]
-      to this end use the FTLR. To do so, we need to show
-      \begin{itemize}
-        \item $\npair[n']{(\start,\addrend)} \in \readCond{}(\local)(W')$
-        \item $\npair[n']{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W')$
-      \end{itemize}
-      As we are considering a public future world, some region still contains $\iota^\pwl (\dom(\ms))$, so it is the same argument as \ref{pf:stack-cap-read} and \ref{pf:stack-cap-write}, respectively.
-    \item $\npair{(\rwx,\start,\addrend)} \in \execCond{}(\local)(W)$ : \\
-      Same argument as \ref{pf:stack-cap-exec}, but also use \ref{lem:nwl-subset-pwl}.
-    \item $\npair{(\exec,\start,\addrend)} \in \execCond{}(\local)(W)$
-      Same argument as \ref{pf:stack-cap-exec}.
+      \item $\npair{(\rwlx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+        Using \ref{pf:stack-cap-write} and \ref{pf:stack-cap-read}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
+      \item $\npair{(\rwx,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+        Using \ref{pf:stack-cap-write} and \ref{pf:stack-cap-read}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
+      \item $\npair{(\exec,\start_\stk, \addrend_\stk)} \in \execCond{}(\local)(W)$\\
+        Using \ref{pf:stack-cap-read} and \ref{pf:stack-cap-write}, we can use Lemma~\ref{lem:stack-exec-cond-help}.
   \end{enumproof}

### 2017-01-31 local-capability-machine.tex
Added missing configuration in anti-red O proof.
Some easy cases of the FTLR.

--- 2017-01-31 local-capability-machine.tex
@@
            &   & \fail \mid
-                 \halt \mid 
+                 \halt 
   \end{array}
@@
 \end{align*}
+\dominique{mention that it is a simplification to take decode total?}
 Assume an $\encodePerm$, $\encodeLoc$, and $\encodePermPair$ function that encodes a permissions, locality, and permission pair, respectively, as an integer:
@@
 
+\subsubsection{LR Sanity lemmas}
+
+\begin{lemma}[Heap satisfaction downwards closure]
+  \label{heap-sat-dc}
+  \begin{align*}
+    &\forall \hs, n' \leq n, W \ldotp \\
+    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
+  \end{align*}
+\end{lemma}
+
 \subsubsection{Fundamental theorem of logical relations}
@@
     holds for all $n' < n$.
-  \item The case that $i = 0$ is impossible, since
+  \item $i \neq 0$, since
     $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. Assume w.l.o.g. that $i = 1+i'$
+    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. \\
+    Therefore, assume w.l.o.g. that $i = 1+i'$,
+    \begin{equation*}
+      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step \var{conf'} \step[i'] (\halted,\Phi)
+    \end{equation*}
+  \item $\start \leq \addr \leq \addrend$, $\perm \in \{ \exec,\rwx, \rwlx \}$
+    and $\var{conf'} =
+    \sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+    \uplus \ms_f))$\\
+    By definition of
+    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)
+    \step \var{conf'}$, using the fact that $\var{conf'} \neq \failed$ (since it
+    is not possible that $\failed \step[i'] (\halted,\Phi)$)
+  \item We do a case analysis on
+    $\memheap(a)$
+    \begin{enumproof}
+    \item Case $\memheap(a) = \jmp{\lv}$:
+    \item Case $\memheap(a) = \jnz{\lv}{\rv}$:
+    \item Case $\memheap(a) = \move{\lv}{\rv}$:
+    \item Case $\memheap(a) = \load{\lv}{\hv}$:
+    \item Case $\memheap(a) = \store{\hv}{\rv}$:
+    \item Case $\memheap(a) = \plus{\lv}{\rv}{\rv}$:
+    \item Case $\memheap(a) = \lea{\lv}{\rv}$:
+    \item Case $\memheap(a) = \restricttwo{\lv}{\rv}$:
+    \item Case $\memheap(a) = \subseg{\lv}{\rv}{\rv}$:
+    \item Case $\memheap(a) = \isptr{\lv}{\rv}$:
+    \item Case $\memheap(a) = \getp{\lv}{\lv}$:
+    \item Case $\memheap(a) = \getl{\lv}{\lv}$:
+    \item Case $\memheap(a) = \getb{\lv}{\lv}$:
+    \item Case $\memheap(a) = \gete{\lv}{\lv}$:
+    \item Case $\memheap(a) = \geta{\lv}{\lv}$:
+    \item Case $\memheap(a) = \fail$: Contradiction, since then
+      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+      \uplus \ms_f)) = \failed$, so we have that $\failed \step[i']
+      (\halted,\Phi)$ which is not possible.
+    \item Case $\memheap(a) = \halt$:
+    \end{enumproof}
+    \begin{enumproof}
     \item
+      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+      \uplus \ms_f)) = (\halted, (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+      \uplus \ms_f))$\\
+      By definition of $\sem{\cdot}$
+    \item Then $i' = 0$ and $\Phi =
+      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+      \ms_f)$. Follows from $(\halted,
+      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+      \ms_f)) \step[i'] (\halted,\Phi)$
+    \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
+      $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and
+      $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{heap-sat-dc}).
+    \end{enumproof}
   \end{enumproof}
@@
 We expect the following lemmas to hold true:
-\begin{lemma}[Downwards closure]
-  \begin{align*}
-    &\forall \hs, n' \leq n, W \ldotp \\
-    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
-  \end{align*}
-\end{lemma}
 

### 2017-01-31 local-capability-machine.tex
Oops, fix latex bug.
The rest of the scall proof

--- 2017-01-31 local-capability-machine.tex
@@
     & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
-    & \quad \npair[n-i]{(\reg',\ms')} \in \observations(W') \\
+    & \quad \npair[n']{(\reg',\ms')} \in \observations(W') \\
     & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
@@
     \item $W_2 \futurestr W_1$ \label{ass:anti-red-ftrworld}
-    \item $(\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f))$ \label{ass:anti-red-exec}
+    \item $\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)$ \label{ass:anti-red-exec}
     \item $\npair[n']{(\reg',\ms')} \in \observations(W_2)$ \label{ass:anti-red-obs}
@@
     \label{ass:anti-red-part-exec}
-    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[i] 
+    (\reg,\ms \uplus \ms_r \uplus \ms_{\var{frame}}) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_{\var{frame}})
   \end{equation}
@@
     \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus \ms_\act \uplus \ms_f),\iota^\sta(\temp,\ms_{\mathit{unused}}'')]$
+    \item $W'' \futurewk \revokeTemp{W}$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
           = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
-          \in \stdvr(W'') &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\} %revokeTemp here?
+          \in \stdvr(\revokeTemp{W''}) &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\}
         \end{cases}
@@
       Use Lemma~\ref{lem:stack-cap-vr} with \ref{pf:scall-works:item:points-to-empty-and-unusedp} and that $W_1$ has region $\iota^{\pwl}(\dom(\ms_{\mathit{unused}}')$.
-    \item $\npair[n-1]{c_{\var{ret}}} \in \stdvr(W_1)$
-      % TODO
+    \item $\npair[n-1]{c_{\var{ret}}} \in \stdvr(W_1)$ \\
+      To this end let 
+      \begin{enumproof}
+        \item $n' < n-1$ \label{pf:scall-works:np-n1}
+        \item $W_2 \futurewk W_1$ \label{pf:scall-works:w2-futurewk-w1}
+      \end{enumproof}
+      be given and show
+      \[
+        \npair[n']{\updatePcPerm{c_{\var{ret}}}} \in \stder(W_2)
+      \]
+      To this assume 
+      \begin{enumproof}[resume]
+        \item $n'' \leq n'$
+        \item $\npair[n'']{\reg_2} \in \stdrr(W_2)$ \label{pf:scall-worsk:reg2-reg-rel}
+        \item $\memSat[n'']{\ms'}{W_2}$ \label{pf:scall-works:w2-memsat}
+      \end{enumproof}
+      be given and show
+      \begin{equation}
+        \label{pf:scall-works:obsw2}
+        \npair[n'']{(\reg_2\update{\pcreg}{\updatePcPerm{c_{\var{ret}}}},\ms')} \in \observations(W_2)
+      \end{equation}
+
+      From \ref{pf:scall-works:w2-futurewk-w1} and \ref{pf:scall-works:w2-memsat}, we can deduce that the memory can be split in the following way:
+      \[
+        \ms' = \ms'' \uplus \ms_r \uplus \ms_\stk \uplus \ms_\act \uplus \ms_\unused'' \uplus \ms_f
+      \]
+      where $\ms''$ is the "permanent" part of memory we get from Lemma~\ref{lem:priv-mono-like}, $\ms_r$ is the part "revoked" of memory from the same lemma that is not otherwise specified, and $\dom(\ms_\unused') = \dom(\ms_\unused'')$.
+From Lemma~\ref{lem:priv-mono-like} we also get
+      \begin{enumproof}[resume]
+        \item $\memSat[n'']{\ms''}{\revokeTemp{W_2}}$ \label{pf:scall-works:rt-w2-memsat}
+      \end{enumproof}
+
+      Assume $n''$ is large enough to execute the rest of the \texttt{scall} instructions.
+If $n''$ is not large enough, then \ref{pf:scall-works:obsw2} is trivial to show.
+To show \ref{pf:scall-works:obsw2} apply Lemma~\ref{lem:anti-red-obs} again where $\ms_r$ is the revoked part.
+Let $\ms_{\var{frame}}'$ be given, the execution until just after the \texttt{scall} proceeds as follows:
+      \[
+        (\reg_2\update{\pcreg}{\updatePcPerm{c_{\var{ret}}}},\ms' \uplus \ms_{var{frame}}') \step[j] (\reg_3,\ms' \uplus \ms_{\var{frame}}')
+      \]
+      where
+      \begin{enumproof}[resume]
+      \item \label{pf:scall-works:reg3-contents}
+        \[
+          \reg_3(r) =
+          \begin{cases}
+            c_{\var{next}} & r = \pcreg \\
+            c_\stk & r = r_\stk \\
+            \reg(r) & r \in \{\overline{r_{\var{priv}}} \}\\
+            \reg_2(r) & \text{ otherwise}
+          \end{cases}
+        \]
+        \item \pointstostack{\reg_3}{\ms_\stk}{\ms_\act \uplus \ms_\unused''} \label{pf:scall-works:reg3-stack}
+      \end{enumproof}
+
+      At this point, we use Hyp-Cont (\ref{pf:scall-works:item:Hyp-Cont}) to show the observation predicate condition of Lemma~\ref{lem:anti-red-obs}:
+      \[
+        \npair[n'']{(reg_3,\ms'' \uplus \ms_\stk \uplus \ms_\act \uplus \ms_\unused'' \uplus \ms_f)} \in \observations(W_2)
+      \]
+      which 
+      \begin{itemize}
+      \item $n'' \leq n-2$ \\
+        Follows from (\ref{pf:scall-works:np-n1})
+      \item $W_2 \futurewk \revokeTemp{W}$ \\
+        We have 
+        \[
+          W_1 \futurewk \revokeTemp{W}
+        \]
+        and assumption \ref{pf:scall-works:w2-futurewk-w1} we get this by transitivity of $\futurewk$.
+      \item $\memSat[n'']{\ms''}{\revokeTemp{W_2}}$ \\
+        Exactly \ref{pf:scall-works:rt-w2-memsat}.        
+      \item for all $r$, we have that:
+        \begin{equation*}
+          \reg_3(r)
+          \begin{cases}
+            = c_{\mathit{next}} &\text{ if } r = \pcreg\\
+            = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
+            \in \stdvr(\revokeTemp{W_2}) &\text{ if $\reg_3(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\} %revokeTemp here?
+          \end{cases}
+        \end{equation*}\\
+        The two first cases follows from \ref{pf:scall-works:reg3-contents}. The third follow from assumption \ref{pf:scall-worsk:reg2-reg-rel} and \ref{lem:stdvr-glob-priv-mono}.
+      \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_\act \uplus \ms_{\mathit{unused}}''$ unused}$\\
+        Exactly \ref{pf:scall-works:reg3-stack}.
+      \end{itemize}
   \end{enumproof}

### 2017-01-31 local-capability-machine.tex

@@
     holds for all $n' < n$.
-  \item The case that $i = 0$ is impossible, since
+  \item $i \neq 0$, since
     $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. Assume w.l.o.g. that $i = 1+i'$
+    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. \\
+    Therefore, assume w.l.o.g. that $i = 1+i'$,
+    \begin{equation*}
+      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step \var{conf'} \step[i'] (\halted,\Phi)
+    \end{equation*}
+  \item $\start \leq \addr \leq \addrend$, $\perm \in \{ \exec,\rwx, \rwlx \}$
+    and $\var{conf'} =
+    \sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+    \uplus \ms_f))$\\
+    By definition of
+    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)
+    \step \var{conf'}$, using the fact that $\var{conf'} \neq \failed$ (since it
+    is not possible that $\failed \step[i'] (\halted,\Phi)$)
+  \item We do a case analysis on
+    $\memheap(a)$
+    \begin{enumproof}
+    \item Case $\memheap(a) = \jmp{\lv}$:
+    \item Case $\memheap(a) = \jnz{\lv}{\rv}$:
+    \item Case $\memheap(a) = \move{\lv}{\rv}$:
+    \item Case $\memheap(a) = \load{\lv}{\hv}$:
+    \item Case $\memheap(a) = \store{\hv}{\rv}$:
+    \item Case $\memheap(a) = \plus{\lv}{\rv}{\rv}$:
+    \item Case $\memheap(a) = \lea{\lv}{\rv}$:
+    \item Case $\memheap(a) = \restricttwo{\lv}{\rv}$:
+    \item Case $\memheap(a) = \subseg{\lv}{\rv}{\rv}$:
+    \item Case $\memheap(a) = \isptr{\lv}{\rv}$:
+    \item Case $\memheap(a) = \getp{\lv}{\lv}$:
+    \item Case $\memheap(a) = \getl{\lv}{\lv}$:
+    \item Case $\memheap(a) = \getb{\lv}{\lv}$:
+    \item Case $\memheap(a) = \gete{\lv}{\lv}$:
+    \item Case $\memheap(a) = \geta{\lv}{\lv}$:
+    \item Case $\memheap(a) = \fail$: Contradiction, since then
+      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+      \uplus \ms_f)) = \failed$, so we have that $\failed \step[i']
+      (\halted,\Phi)$ which is not possible.
+    \item Case $\memheap(a) = \halt$:
+      \begin{enumproof}
       \item
+        $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+        \uplus \ms_f)) = (\halted, (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+        \uplus \ms_f))$\\
+        By definition of $\sem{\cdot}$
+      \item Then $i' = 0$ and $\Phi =
+        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+        \ms_f)$. Follows from $(\halted,
+        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+        \ms_f)) \step[i'] (\halted,\Phi)$
+      \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
+        $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and
+        $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{heap-sat-dc}).
+      \end{enumproof}
+    \end{enumproof}
   \end{enumproof}

### 2017-01-31 local-capability-machine.tex
put old incomplete awkward example proof in comment
added lemma for memory satisfaction of static region
added lemma about data only memory and memory satisfaction for standard regions.

@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using scall)]
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using \texttt{scall} lemma)]
   Let $n$ be given and make the assumptions of the lemma. Define
@@
 
+\begin{comment}
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-g1}]
@@
 \lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
+\end{comment}
 
@@
 
+\begin{lemma}[Memory satisfaction and static regions]
+  \label{lem:mem-sat-static}
+  \[
+    \memSat{\ms}{[i \mapsto \iota^\sta(v,\ms)]}
+  \]
+\end{lemma}
+
+\begin{lemma}[Data only memory and standard regions]
+  \label{lem:mem-sat-data-only-std-regions}
+  If
+  \begin{itemize}
+  \item $\forall \addr \in \dom(\ms)\ldotp \ms(a) \in \nats$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{\nwl,p} \}$
+  \end{itemize}
+  then
+  \[
+    \memSat{\ms}{[i \mapsto \iota^\pwl(\dom(\ms))]}
+  \]
+\end{lemma}
+
 \subsubsection{Future worlds}

### 2017-01-31 local-capability-machine.tex
Added item to todo. Found issue in awkward example proof, solved it introducing new region.

@@
               In the specification above $\iota_\malloc'$ is a future region of the initial region that governs malloc.
+              \lau{What about $c_\malloc$ and the value relation? Do we add this to the specification or try to prove it based on the specification?}
 
@@
   \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
+  \item $\forall a \in \dom(\ms_\stk) \ldotp \ms_\stk(a) = 0$ %This condition could be weakened to \memSat{\ms_\stk}{[\iota^\pwl(\dom(\ms_\stk))]}
   \end{itemize}
@@
 \end{lemma}
+%
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using \texttt{scall} lemma)]
@@
   \begin{align*}
-    W = & [0 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]\\
-        & [1 \mapsto \iota^\sta (\perma,\ms_\link)]\\
+    W = & [0 \mapsto \iota_{\malloc,0}]\\
+        & [1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)]\\
         & [2 \mapsto \iota^\pwl_{\start_\stk, \addrend_\stk}]\\
         & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
-        & [4 \mapsto \iota_{\malloc,0}]
+        & [4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]
   \end{align*}
@@
   \end{equation}
-  then we are done. The termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 1 is permanent, so
+  then, the termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 4 is permanent, so
   \[
@@
   \]
-  So we need to show \ref{eq:f4:start-conf-in-obs}. 
+  So it suffices to show \ref{eq:f4:start-conf-in-obs}. 
   To this end use the FTLR to show $\npair{c_\adv} \in \stder(W)$, so show
@@
     \item $\npair{(\start_\adv,\addrend_\adv)} \in \readCond{}(\glob)(W)$\\
-      %TODO
+      Simple using region $3$ in $W$.
     \item $\npair{(\start_\adv,\addrend_\adv)} \in \writeCond{}(\iota^\nwl,\glob)(W)$\\
-      %TODO
+      Simple using region $3$ in $W$.
   \end{enumproof}
@@
     \item $\memSat{\ms \uplus \ms_\stk}{W}$\\
-      %TODO
+      \begin{enumproof}
+        \item $\memSat{\ms_{g1} \uplus \ms_\flag}{[4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]}$ \\
+          Lemma~\ref{lem:mem-sat-static}.
+        \item $\memSat{\ms_\stk}{[2 \mapsto \iota^\pwl_{\start_\stk,\addrend_\stk}]}$\\
+          Lemma~\ref{lem:mem-sat-data-only-std-regions} and assumption that $\ms_\stk$ is all 0.
+        \item $\memSat{\ms_{\malloc} \uplus \ms_\link \uplus \ms_\adv}{[0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]}$\\
+          For convenience define
+          \[
+            W_{\var{mini}} = [0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
+          \]
+          Partitioning the memory segment in the components of the disjoint union, and given $n' < n$, the $\malloc$ part follows from assumption $\memSat{\ms_\malloc}{[0 \mapsto \iota_{\malloc,0}]}$ and the $\malloc$ specification.
+
+          The memory link amounts to showing:
+          \[
+            \npair[n']{\ms_\link} \in H^{\sta,u}(\ms_\link)(1)(\xi^{-1}(W_{\var{mini}}))
+          \]
+          which in turn amounts to showing
+          \[
+            \npair[n'-1] \in \stdvr(W_{\var{mini}})
+          \]
+          which is indeed the case. \lau{We need to add a reason here! Is this part of the specification of malloc?}
+
+          Showing
+          \[
+            \npair[n']{\ms_\adv} \in H^\sta_{\start_\adv,\addrend_\adv}(1)(\xi^{-1}(W_{\var{mini}}))
+          \]
+          is a bit more involved. It amounts to 
+          \[
+            \forall \addr \in \dom(\ms_\adv) \ldotp \npair[n'-1]{\ms_\adv(\addr)} \in \stdvr(W_{\var{mini}})
+          \]
+          which in turn is trivial for everything but
+          \[
+            \npair[n'-1]{c_\link} \in \stdvr(W_{\var{mini}})
+          \]
+          This amounts to showing
+          \[
+            \npair[n'-1]{(\link,\link)} \in \readCond{}(\glob)(W_{\var{mini}})
+          \]
+          which amounts to
+          \[
+            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n'-2] \iota^\pwl_{\link,\link}
+          \]
+          which follows from Lemma~\ref{lem:stau-subset-pwl}          
+      \end{enumproof}
+      Using Lemma~\ref{lem:disj-mem-sat} repeatedly with the above gives the desired memory satisfaction.
+
     \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$\\
@@
 \begin{mathpar}
-  \inferrule{(s,H) = (s',H) \\
+  \inferrule{(s,\phi_\pub,\phi) = (s',\phi_\pub',\phi') \\
     \forall \hat{W} \ldotp H \; s \; \hat{W} \nsubeq H' \; s' \; \hat{W} }
-  { (v,s,\phi,H) \nsubsim (v',s',\phi',H')}
+  { (v,s,\phi_\pub,\phi,H) \nsubsim (v',s',\phi_\pub',\phi',H')}
 \end{mathpar}
@@
 
+Static region used for parts of memory that should not change and where you pass control to untrusted code.
+\begin{align*}
+  \iota^{\sta,u} (v,\ms) &= (v,1,=,=,H^{\sta,u}\;\ms)\\
+  H^\sta \; \ms \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
+\end{align*}
+
 \begin{align*}
   \iota^\cnst (v,n) &= (v,1,=,=,H^\cnst\;n)\\
-  H^\cnst \; n \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \Rightarrow \forall a \in \dom(\ms)\ldotp \ms(a) = n \}
+  H^\cnst \; n \; s \; \hat{W} = \left\{\npair{\ms'} \middle|
+    \begin{aligned}
+      &\ms' = \ms \\
+      &\forall \addr \in \dom(\ms) \ldotp \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\}
 \end{align*}
@@
   \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \iota^\nwl_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
+    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp  \\
+    & \quad \iota^\nwl_{\start,\addrend} \nsubsim \iota^\pwl_{\start,\addrend}
   \end{align*}
@@
   \begin{align*}
+    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp  \\
+    & \quad \iota^{\nwl,p}_{\start,\addrend}  \nsubsim \iota^\pwl_{\start,\addrend}
+  \end{align*}
+\end{lemma}
+\begin{lemma}
+  \label{lem:stau-subset-pwl}
+  \begin{align*}
     & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \iota^{\nwl,p}_{\start,\addrend} (v) \nsubsim \iota^\pwl_{\start,\addrend}(v)
+    & \quad \dom(\ms) = [\star,\addrend] \Rightarrow \\
+    & \qquad \iota^{\sta,u}_{\start,\addrend}(v,\ms) \nsubsim \iota^\pwl_{\start,\addrend}
   \end{align*}

### 2017-01-31 local-capability-machine.tex
fixed some bugs in the new regions added

@@
     \item $\npair{(\start_\adv,\addrend_\adv)} \in \readCond{}(\glob)(W)$\\
-      Simple using region $3$ in $W$.
+      Simple using region $3$ in $W$ and Lemma~\ref{lem:nwlp-subset-pwl}.
     \item $\npair{(\start_\adv,\addrend_\adv)} \in \writeCond{}(\iota^\nwl,\glob)(W)$\\
@@
           \]
-          which follows from Lemma~\ref{lem:stau-subset-pwl}          
+          which follows from Lemma~\ref{lem:stau-subset-pwl}.
       \end{enumproof}
       Using Lemma~\ref{lem:disj-mem-sat} repeatedly with the above gives the desired memory satisfaction.
-
     \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$\\
@@
       \begin{enumproof}
-        \item $\npair{c_\stk} \in \stdvr(W)$ \\
-          % TODO
+        \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}(c_\stk)} \in \stdvr(W)$ \\
+          The assumptions on $c_\stk$ and $\ms_\stk$ in the lemma entail
+          \begin{itemize}
+          \item \pointstostack{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}}{\emptyset}{\ms_\stk}
+          \end{itemize}
+          and further there is a $\iota^\pwl$ region for $\ms_\stk$ in $W$, so the result follows from Lemma~\ref{stack-cap-vr}.
         \item $\npair{c_{g1}} \in \stdvr(W)$ \\
@@
   \iota^{\sta,u} (v,\ms) &= (v,1,=,=,H^{\sta,u}\;\ms)\\
-  H^\sta \; \ms \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
-\end{align*}
-
-\begin{align*}
-  \iota^\cnst (v,n) &= (v,1,=,=,H^\cnst\;n)\\
-  H^\cnst \; n \; s \; \hat{W} = \left\{\npair{\ms'} \middle|
+  H^\sta \; \ms \; s \; \hat{W} = & \left\{\npair{\ms'} \middle|
     \begin{aligned}
       &\ms' = \ms \\
-      &\forall \addr \in \dom(\ms) \ldotp \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))
+      &\forall \addr \in \dom(\ms) \ldotp \npair[n-1]{\ms(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
     \end{aligned}
         \right\}
+\end{align*}
+
+\begin{align*}
+  \iota^\cnst (v,n) &= (v,1,=,=,H^\cnst\;n)\\
+  H^\cnst \; n \; s \; \hat{W} = &  \{\npair{\ms} \mid n > 0 \land \forall \addr \in \dom(\ms) \ldotp \ms(\addr) = n \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
 \end{align*}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:nwl-subset-pwl}]
+
+\end{proof}
+
 \begin{lemma}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:nwlp-subset-pwl}]
+
+\end{proof}
+
 \begin{lemma}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stau-subset-pwl}]
+
+\end{proof}
 

### 2017-01-31 local-capability-machine.tex
added some helper lemmas to help prove standard region lemmas
added more lemmas about the permission based conditions

@@
 \begin{lemma}
+  If
+  \[
+    \npair{\stdcap} \in \stdvr(\revokeTemp{W})
+  \]
+  then 
+  \[
+    \gl = \glob
+  \]
+\end{lemma}
+\begin{lemma}
+  If
+  \[
+    \npair{w} \in \stdvr(\revokeTemp{W})
+  \]
+  then
+  \[
+    \npair{w} \in \stdvr(W)
+  \]
+\end{lemma}
+
+\begin{lemma}
   \label{lem:nwl-subset-pwl}
@@
   \item $W' \futurewk W$
-  \item $\npair{\start,\addrend} \in \readCond{}(\gl)(W)$
+  \item $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
   \end{itemize}
@@
   \[
-    \npair{\start,\addrend} \in \readCond{}(\gl)(W')
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W')
   \]
@@
 
+\begin{lemma}[$\mathit{readCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:readcond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\npair{(\start,\addrend)} \in \readCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-priv}]
+  % TODO
+\end{proof}
+
 \begin{lemma}[$\mathit{readCondition}$ downwards-closed]
@@
   \item $ n' \leq n$
-  \item $\npair[n]{\start,\addrend} \in \readCond{}(\gl)(W)$
+  \item $\npair[n]{(\start,\addrend)} \in \readCond{}(\gl)(W)$
   \end{itemize}
@@
   \[
-    \npair[n']{\start,\addrend} \in \readCond{}(\gl)(W)
+    \npair[n']{(\start,\addrend)} \in \readCond{}(\gl)(W)
   \]
@@
   \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
   \end{itemize}
@@
   \[
-    \npair{\start,\addrend}\in \writeCond{}(\iota,\gl)(W')
+    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\gl)(W')
   \]
@@
 
+\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:writecond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{\(start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
+  % TODO
+\end{proof}
+
 \begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
@@
   \[
-    \npair[n']{\start,\addrend} \in \writeCond{}(\iota,\gl)(W)
+    \npair[n']{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
   \]
@@
 \begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
+  % TODO
+\end{proof}
+
+%ExecCond
+\begin{lemma}[$\mathit{execCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:execcond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\perm \in \{\exec, \rwx, \rwlx\}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)}\in \execCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{execCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:execcond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\perm \in \{\exec, \rwx \}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)}\in \execCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{execCondition}$ downwards-closed]
+  \label{lem:execcond-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\perm \in \{\exec, \rwx, \rwlx\}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-dc}]
+  % TODO
+\end{proof}
+
+% EnterCond
+\begin{lemma}[$\mathit{enterCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:entrycond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{enterCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:entrycond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{enterCondition}$ downwards-closed]
+  \label{lem:entrycond-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-dc}]
   % TODO
@@
   \begin{align*}
-    & \forall n, \perm, \start, \addrend, \addr, W \ldotp \\
-    & \quad  \npair{\stdcap[(\perm,\glob)]} \in \stdvr(W) \\
-    & \qquad \Rightarrow \npair{\stdcap[(\perm,\glob)]} \in \stdvr(\wrev{W}) 
+    & \forall n, \perm, \start, \addrend, \addr, W, W' \ldotp \\
+    & \quad  \npair{\stdcap[(\perm,\glob)]} \in \stdvr(W) \land W' \futurestr W\\
+    & \qquad \Rightarrow \npair{\stdcap[(\perm,\glob)]} \in \stdvr(W') 
   \end{align*}
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stdvr-glob-priv-mono}]
+  Assume 
+  \begin{enumproof}
+  \item $W' \futurestr W$
+  \item $\npair{\stdcap[(\perm,\glob)]} \in \stdvr(W)$
+  \end{enumproof}
+  and show
+  \[
+    \npair{\stdcap[(\perm,\glob)]} \in \stdvr(W')
+  \]
+  to this end consider all the possible cases of $\perm$:
+  \begin{enumproof}
+    \item $\perm = \noperm$\\
+      Trivial
+    \item $\perm = \readonly$ \\
+
+    \item $\perm = \readwrite$ \\
+
+    \item $\perm = \rwl$ \\
+
+    \item $\perm = \exec$ \\
+
+    \item $\perm = \rwx$ \\
+
+    \item $\perm = \rwlx$ \\
+  \end{enumproof}
+
+\end{proof}
 

### 2017-01-31 local-capability-machine.tex
typo
Some work on FTLR proof.  Move some more lemmas forward.

@@
 
-%% make latex preview-mode work with natbib...
-%\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
+%make latex preview-mode work with natbib...
+\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
 
@@
 
+\subsubsection{Downward closedness lemmas}
+\label{sec:downward-closedness}
+
+\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
+  \label{lem:readcond-dc}
+  If
+  \begin{itemize}
+  \item $ n' \leq n$
+  \item $\npair[n]{(\start,\addrend)} \in \readCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair[n']{(\start,\addrend)} \in \readCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
+  \label{lem:writecond-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{execCondition}$ downwards-closed]
+  \label{lem:execcond-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\perm \in \{\exec, \rwx, \rwlx\}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{enterCondition}$ downwards-closed]
+  \label{lem:entrycond-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\stdvr{\cdot}$ downwards-closed]
+  \label{lem:stdvr-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\npair{\word} \in \stdvr(W)$
+  \end{itemize}
+  then 
+  \[
+   \npair[n']{\word} \in \stdvr(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stdvr-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\stdrr{\cdot}$ downwards-closed]
+  \label{lem:stdrr-dc}
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\npair{\reg} \in \stdrr(W)$
+  \end{itemize}
+  then 
+  \[
+   \npair[n']{\reg} \in \stdrr(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stdrr-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[Heap satisfaction downwards closure]
+  \label{lem:heap-sat-dc}
+  \begin{align*}
+    &\forall \hs, n' \leq n, W \ldotp \\
+    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
+  \end{align*}
+\end{lemma}
+
+\subsubsection{Monotonicity lemmas}
+\label{sec:mono}
+
+\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:readcond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:readcond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\npair{(\start,\addrend)} \in \readCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-priv}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:writecond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:writecond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
+  % TODO
+\end{proof}
+
+%ExecCond
+\begin{lemma}[$\mathit{execCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:execcond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\perm \in \{\exec, \rwx, \rwlx\}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)}\in \execCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{execCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:execcond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\perm \in \{\exec, \rwx \}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)}\in \execCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
+  % TODO
+\end{proof}
+
+% EnterCond
+\begin{lemma}[$\mathit{enterCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:entrycond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{enterCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:entrycond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
+  % TODO
+\end{proof}
+
 \subsubsection{Antireduction for the observation relation}
@@
 
-\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:readcond-mono-pub}
+\subsubsection{Fundamental theorem of logical relations}
+
+\begin{lemma}[Conditions for load instruction are sufficient]
+  \label{lem:conds-load-suffice}
   If
   \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W')
-  \]
+  \item $\heapSat[\memheap]{n}{W}$
+  \item $\var{c} = ((\perm',\gl'),\start',\addrend',\addr')$
+  \item $\npair{c}\in\stdvr(W)$
+  \item $\readAllowed{\perm'}$ 
+  \item $\withinBounds{\var{c}}$
+  \end{itemize}
+
+  then $\npair{\memheap(\addr')} \in \stdvr(W)$
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
-  % TODO
+\begin{proof}
+  TODO
 \end{proof}
 
-\begin{lemma}[$\mathit{readCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:readcond-mono-priv}
+\begin{lemma}[Conditions for lea instruction are sufficient]
+  \label{lem:conds-lea-suffice}
   If
   \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\npair{(\start,\addrend)} \in \readCond{}(\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\glob)(W')
-  \]
+  \item $\npair{((\perm,\gl),\start,\addrend,\addr)}\in\stdvr(W)$
+  \item $\perm \neq \entry$
+  \end{itemize}
+
+  then $\npair{((\perm,\gl),\start,\addrend,\addr')} \in \stdvr(W)$
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-priv}]
-  % TODO
+\begin{proof}
+  Follows by inspection of the cases in the definition of $\stdvr(W)$: $a$ is
+  ignored in all cases except where $\perm = \entry$.
 \end{proof}
 
-\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
-  \label{lem:readcond-dc}
+\begin{lemma}[Conditions for restrict instruction are sufficient]
+  \label{lem:conds-restrict-suffice}
   If
   \begin{itemize}
-  \item $ n' \leq n$
-  \item $\npair[n]{(\start,\addrend)} \in \readCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair[n']{(\start,\addrend)} \in \readCond{}(\gl)(W)
-  \]
+  \item $\npair{((\perm,\gl),\start,\addrend,\addr)}\in\stdvr(W)$
+  \item $(\perm',\gl')\sqsubseteq (\perm,\gl)$
+  \end{itemize}
+
+  then  $\npair{((\perm',\gl'),\start,\addrend,\addr)}\in \stdvr(W)$
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
-  % TODO
+\begin{proof}
+  TODO
 \end{proof}
 
-\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:writecond-mono-pub}
+\begin{lemma}[Conditions for subseg instruction are sufficient]
+  \label{lem:conds-subseg-suffice}
   If
   \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\gl)(W')
-  \]
+  \item $\npair{((\perm',\gl'),\start',\addrend',\addr')} \in \stdvr(W)$
+  \item $\start' \leq \start''$
+  \item $\addrend'' \leq \addrend'$
+  \end{itemize}
+
+  then $\npair{((\perm',\gl'),\start'',\addrend'',\addr')} \in \stdvr(W)$
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
-  % TODO
+\begin{proof}
+  TODO
 \end{proof}
 
-\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:writecond-mono-priv}
+\begin{lemma}[Conditions for store instruction are sufficient]
+  \label{lem:conds-store-suff}
   If 
   \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{\(start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
-  \]
+  \item $\heapSat[m]{n}{W}$
+  \item $((\perm,\gl),\start,\addrend,\addr) = c$
+  \item $\npair{c}\in\stdvr(W)$
+  \item $\writeAllowed{\perm}$
+  \item $\withinBounds{\var{c}}$
+  \item $\npair{\var{w}}\in\stdvr(W)$
+  \item if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in
+    \{\rwlx,\readwritel \}$
+  \end{itemize}
+
+  then $\heapSat[{m\update{\addr}{\var{w}}}]{n}{W}$
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
-  % TODO
+\begin{proof}
+  TODO
 \end{proof}
-
-\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
-  \label{lem:writecond-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
-  \end{itemize}
-  then 
-  \[
-    \npair[n']{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
-  % TODO
-\end{proof}
-
-%ExecCond
-\begin{lemma}[$\mathit{execCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:execcond-mono-pub}
-  If
-  \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\perm \in \{\exec, \rwx, \rwlx\}$
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\perm,\start,\addrend)}\in \execCond{}(\iota,\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{execCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:execcond-mono-priv}
-  If
-  \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\perm \in \{\exec, \rwx \}$
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\perm,\start,\addrend)}\in \execCond{}(\glob)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{execCondition}$ downwards-closed]
-  \label{lem:execcond-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\perm \in \{\exec, \rwx, \rwlx\}$
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
-  \end{itemize}
-  then 
-  \[
-    \npair[n']{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-dc}]
-  % TODO
-\end{proof}
-
-% EnterCond
-\begin{lemma}[$\mathit{enterCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:entrycond-mono-pub}
-  If
-  \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\iota,\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{enterCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:entrycond-mono-priv}
-  If
-  \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\glob)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{enterCondition}$ downwards-closed]
-  \label{lem:entrycond-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
-  \end{itemize}
-  then 
-  \[
-    \npair[n']{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:entrycond-dc}]
-  % TODO
-\end{proof}
-
-\subsubsection{LR Sanity lemmas}
-
-\begin{lemma}[Heap satisfaction downwards closure]
-  \label{heap-sat-dc}
-  \begin{align*}
-    &\forall \hs, n' \leq n, W \ldotp \\
-    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
-  \end{align*}
-\end{lemma}
-
-\subsubsection{Fundamental theorem of logical relations}
 
@@
   and
-  \begin{itemize}
+  \begin{enumproof}
   \item $\npair[n]{\reg}\in \stdrr(W)$
@@
   \item $i \leq n$
-  \item $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$
-  \end{itemize}
+  \item $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)$
+  \item $\Phi \step[i] (\halted,\Phi')$
+  \end{enumproof}
 
   Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-  $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
+  $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
 \end{lemma}
@@
     $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. \\
+    \ms_f)\neq (\halted,\Phi')$ for any $\Phi'$. \\
     Therefore, assume w.l.o.g. that $i = 1+i'$,
     \begin{equation*}
-      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step \var{conf'} \step[i'] (\halted,\Phi)
+      \Phi \step \var{conf'} \step[i'] (\halted,\Phi')
     \end{equation*}
-  \item $\start \leq \addr \leq \addrend$, $\perm \in \{ \exec,\rwx, \rwlx \}$
-    and $\var{conf'} =
-    \sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-    \uplus \ms_f))$\\
-    By definition of
-    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)
-    \step \var{conf'}$, using the fact that $\var{conf'} \neq \failed$ (since it
-    is not possible that $\failed \step[i'] (\halted,\Phi)$)
-  \item We do a case analysis on
-    $\memheap(a)$
-    \begin{enumproof}
-    \item Case $\memheap(a) = \jmp{\lv}$:
-    \item Case $\memheap(a) = \jnz{\lv}{\rv}$:
-    \item Case $\memheap(a) = \move{\lv}{\rv}$:
-    \item Case $\memheap(a) = \load{\lv}{\hv}$:
-    \item Case $\memheap(a) = \store{\hv}{\rv}$:
-    \item Case $\memheap(a) = \plus{\lv}{\rv}{\rv}$:
-    \item Case $\memheap(a) = \lea{\lv}{\rv}$:
-    \item Case $\memheap(a) = \restricttwo{\lv}{\rv}$:
-    \item Case $\memheap(a) = \subseg{\lv}{\rv}{\rv}$:
-    \item Case $\memheap(a) = \isptr{\lv}{\rv}$:
-    \item Case $\memheap(a) = \getp{\lv}{\lv}$:
-    \item Case $\memheap(a) = \getl{\lv}{\lv}$:
-    \item Case $\memheap(a) = \getb{\lv}{\lv}$:
-    \item Case $\memheap(a) = \gete{\lv}{\lv}$:
-    \item Case $\memheap(a) = \geta{\lv}{\lv}$:
-    \item Case $\memheap(a) = \fail$: Contradiction, since then
-      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-      \uplus \ms_f)) = \failed$, so we have that $\failed \step[i']
-      (\halted,\Phi)$ which is not possible.
-    \item Case $\memheap(a) = \halt$:
-      \begin{enumproof}
-      \item
-        $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-        \uplus \ms_f)) = (\halted, (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-        \uplus \ms_f))$\\
-        By definition of $\sem{\cdot}$
-      \item Then $i' = 0$ and $\Phi =
-        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-        \ms_f)$. Follows from $(\halted,
+  \item $n > 0$, since otherwise $i = 0$ (because $i \leq n$) and this is
+    impossible by the previous point.
+  \item By inspection of the definitions of $\Phi\step \var{conf'}$ and
+    $\sem{\decode(\memheap(\addr))}$ and $\updatePcPerm{\cdot}$ and
+    $\stdUpdatePc{\cdot}$, it is easy to see that one of the following cases
+    must hold:
+  \item Case $\var{conf'} = \failed$: contradiction, since it is not possible
+    that $\failed \step[i'] (\halted,\Phi)$. 
+  \item Case $\var{conf'} = (\halted,\Phi)$:
+    \begin{enumproof}
+    \item Then $i' = 0$ and $\Phi' =
+      \Phi$\\
+      Follows from $(\halted,
       (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
@@
     \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
-        $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and
-        $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{heap-sat-dc}).
-      \end{enumproof}
+      $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and
+      $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
+    \end{enumproof}
+  \item Case $\var{conf'} = \Phi'$, and additionally, one of the following
+    holds:
+    \begin{itemize}
+    \item $\memheap[\Phi'] = \memheap[\Phi]$
+    \item $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$, with
+      $\memreg[\Phi](r_1) = ((\perm',\gl'),\start',\addrend',\addr') = c$ and
+      $\writeAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
+      \memreg(r_2)$ and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm' \in
+      \{\rwlx,\readwritel \}$
+    \end{itemize}
+    and also one of the following holds:
+    \begin{itemize}
+    \item $\memreg[\Phi'](\pcreg) = \updatePcPerm{\memreg(\lv)}$
+    \item $\memreg[\Phi'](\pcreg) = \var{newPc}$ and $\var{newPc} =
+      ((\perm,\gl),\start,\addrend,\addr + 1)$
+      \dominique{This is wrong: the semantics allow, for example, a move into
+        the PC, after which the new pc is increased by one}
+    \end{itemize}
+    and finally, for all $r \in (\RegName \setminus \{\pcreg\})$, one of the following
+    holds:
+    \begin{itemize}
+    \item $\memreg[\Phi'](r) = \memreg[\Phi](r)$
+    \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$
+    \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =
+      ((\perm',\gl'),\start',\addrend',\addr') = \var{c}$ and
+      $\readAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
+      \memheap(\addr')$
+    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
+      ((\perm',\gl'),\start',\addrend',\addr')$ and $\perm' \neq \entry$ and
+      $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in
+      \ints$
+    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+      ((\perm',\gl'),\start',\addrend',\addr')$ and $(\perm'',\gl'')\sqsubseteq
+      (\perm',\gl')$ and $c = ((\perm'',\gl''),\start',\addrend',\addr')$
+    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+      ((\perm',\gl'),\start',\addrend',\addr')$ and $\start' \leq \start''$ and
+      $\addrend'' \leq \addrend'$ and $c =
+      ((\perm',\gl'),\start'',\addrend'',\addr')$
+    \end{itemize}
+    In this case, we have:
+    \begin{enumproof}
+    \item $\heapSat[{\memheap[\Phi']}]{n-1}{W}$.
+      \begin{enumproof}
+      \item Case $\memheap[\Phi'] = \memheap[\Phi]$:
+        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
+      \item Case $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$,
+        with $\memreg[\Phi](r_1) = ((\perm',\gl'),\start',\addrend',\addr') = c$
+        and $\writeAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
+        \memreg(r_2)$ and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm'
+        \in \{\rwlx,\readwritel \}$.\\
+        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by
+        Lemmas~\ref{lem:conds-store-suff} and~\ref{lem:heap-sat-dc} using the
+        fact that $\heapSat[\memheap]{n}{W}$ and
+        $\npair{\memreg(r_1)}\in\stdvr(W)$ and
+        $\npair{\memreg(r_2)}\in\stdvr(W)$ which follows from
+        $\npair{\memreg}\in\stdrr(W)$ by definition.
+      \end{enumproof}
+    \item $\npair[n-1]{\memreg[\Phi']}\in\stdrr(W)$.
+      \begin{enumproof}
+      \item Assume: $r \in \RegName\setminus\{\pcreg\}$.\\
+        Suffices: $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$.\\
+        By: definition of $\stdrr{\cdot}$
+      \item Case $\memreg[\Phi'](r) = \memreg[\Phi](r)$:
+        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows from
+        $\npair{\memreg}\in\stdrr(W)$ by definition of $\stdrr(\cdot)$ using
+        Lemma~\ref{lem:stdvr-dc}.
+      \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$.
+        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by definition of
+        $\stdvr(\cdot)$
+      \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =
+        ((\perm',\gl'),\start',\addrend',\addr') = \var{c}$ and
+        $\readAllowed{\perm'}$ and $\withinBounds{\var{c}}$ and $\var{w} =
+        \memheap(\addr')$:\\
+        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        Lemmas~\ref{lem:conds-load-suffice} and~\ref{lem:stdvr-dc} using the
+        fact that $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$
+        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
+      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
+        ((\perm',\gl'),\start',\addrend',\addr')$ and $\perm' \neq \entry$ and
+        $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in
+        \ints$: \\
+        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        Lemmas~\ref{lem:conds-lea-suffice} and~\ref{lem:stdvr-dc} using the
+        fact that $\npair{\memreg(r_1)}\in\stdvr(W)$
+        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
+      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+        ((\perm',\gl'),\start',\addrend',\addr')$ and $(\perm'',\gl'')\sqsubseteq
+        (\perm',\gl')$ and $c = ((\perm'',\gl''),\start',\addrend',\addr')$: \\
+        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        Lemmas~\ref{lem:conds-restrict-suffice} and~\ref{lem:stdvr-dc} using the
+        fact that $\npair{\memreg(r)}\in\stdvr(W)$
+        which follows from $\npair{\memreg}\in\stdrr(W)$ by
+        definition\dominique{this does not follow if $r = \pcreg$, similar
+          problem in other cases}.
+      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+        ((\perm',\gl'),\start',\addrend',\addr')$ and $\start' \leq \start''$ and
+        $\addrend'' \leq \addrend'$ and $c =
+        ((\perm',\gl'),\start'',\addrend'',\addr')$: \\
+        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        Lemmas~\ref{lem:conds-subseg-suffice} and~\ref{lem:stdvr-dc} using the
+        fact that $\npair{\memreg(r)}\in\stdvr(W)$
+        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
+      \end{enumproof}
+    \item $\npair[n-1]{\memreg[\Phi'](\pcreg)} \in \stder(W)$: TODO
     \end{enumproof}

### 2017-01-31 local-capability-machine.tex
Some more work on the FTLR.

@@
 
-\begin{lemma}[Fundamental theorem proof helper lemma]
-  \label{lem:ftlr-induction-helper}
+\begin{theorem}[Fundamental theorem of logical relations]
+  \label{thm:ftlr}
   For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
@@
   \end{itemize}
-  and
-  \begin{enumproof}
-  \item $\npair[n]{\reg}\in \stdrr(W)$
-  \item $\heapSat[\hs]{n}{W}$
-  \item $i \leq n$
-  \item $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)$
-  \item $\Phi \step[i] (\halted,\Phi')$
-  \end{enumproof}
-
-  Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-  $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
-\end{lemma}
+  then
+  \[
+    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
+  \]
+\end{theorem}
 \begin{proof}
   \begin{enumproof}
+  \item Assume: $n' \leq n$, $\npair[n']{\reg}\in \stdrr(W)$, $\heapSat[\hs]{n'}{W}$.\\
+    Suffices: $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs)}\in\observations(W)$.\\
+    By: definition of $\stder(W)$.
+  \item Assume: $\ms_f$, $\heap'$, $i \leq n'$, $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$,\\
+    Suffices: $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
+    $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+    By: definition of $\observations(W)$ \label{suff-after-eval}
   \item By induction on $n$. In other words, assume that the theorem already
@@
     impossible by the previous point.
+  \item $\npair{\memreg(\pcreg)}\in\stdvr(W)$. \label{step:ftlr-pc-vr}
+    Proof:
+    \begin{enumproof}
+    \item Assume: $\perm' \in \{\exec,\rwx,\rwlx\}$ with $\perm'
+      \sqsubseteq\perm$\\
+      Suffices: $\npair{(\perm',\start,\addrend)}\in\execCond{g}(W)$\\
+      By: the definition of $\stdvr(\cdot)$ using the assumptions
+    \item Assume: $n'<n$, $W'\future W$, $\addr' \in [\start,\addrend]$, $\gl
+      = \local \Rightarrow \future = \futurewk$,
+      $\gl=\glob\Rightarrow\future = \futurestr$.\\
+      Suffices: $\npair[n']{((\perm,\gl),\start,\addrend,\addr')} \in
+      \stder(W')$.
+      By: definition of $\execCond{g}(W)$
+    \item By induction, using the assumptions and Lemmas~\ref{lem:readcond-dc}
+      and~\ref{lem:writecond-dc}.
+    \end{enumproof}
+  \item For all $r \in \RegName$, $\npair[n]{\memreg(r)}\in\stdvr(W)$. \label{step:ftlr-reg-rr}
+    \begin{enumproof}
+    \item Case $r \neq \pcreg$: follows from $\npair[n]{\reg}\in \stdrr(W)$ by
+      definition of $\stdrr(W)$.
+    \item Case $r = \pcreg$: by step~\ref{step:ftlr-pc-vr}.
+    \end{enumproof}
   \item By inspection of the definitions of $\Phi\step \var{conf'}$ and
@@
         fact that $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$
-        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
+        which we have from step~\ref{step:ftlr-reg-rr}.
       \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
@@
         fact that $\npair{\memreg(r_1)}\in\stdvr(W)$
-        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
+        which we have from step~\ref{step:ftlr-reg-rr}.
       \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
@@
         which follows from $\npair{\memreg}\in\stdrr(W)$ by
-        definition\dominique{this does not follow if $r = \pcreg$, similar
-          problem in other cases}.
+        definition.
       \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
@@
     \end{enumproof}
-  \end{enumproof}
-\end{proof}
-
-\begin{theorem}[Fundamental theorem of logical relations]
-  \label{thm:ftlr}
-  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
-  If one of the following holds:
-  \begin{itemize}
-  \item \[
-      \begin{gathered}
-        \perm = \exec \land\\
-        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-      \end{gathered}
-    \]
-  \item \[
-      \begin{gathered}
-        \perm = \rwx \land \\
-        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
-        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
-      \end{gathered}
-    \]
-  \item \[
-      \begin{gathered}
-        \perm = \rwlx \land\\
-        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
-        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
-      \end{gathered}
-    \]
-  \end{itemize}
-  then
-  \[
-    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
-  \]
-\end{theorem}
-\begin{proof}
-  \begin{enumproof}
-  \item Assume: $n' \leq n$, $\npair[n']{\reg}\in \stdrr(W)$, $\heapSat[\hs]{n'}{W}$.\\
-    Suffices: $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs)}\in\observations(W)$.\\
-    By: definition of $\stder(W)$.
-  \item Assume: $\ms_f$, $\heap'$, $i \leq n'$, $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$,\\
-    Suffices: $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-    $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
-    By: definition of $\observations(W)$ \label{suff-after-eval}
-  \item This follows by Lemma~\ref{lem:ftlr-induction-helper} using
-    Lemmas~\ref{lem:readcond-dc} and~\ref{lem:writecond-dc}.
   \end{enumproof}

### 2017-02-01 local-capability-machine.tex
fixed transitivity issue w/ priv future worlds - temp world can now become anything, not just revoked.
changed priv future world relation (and more)

--- 2017-01-31 local-capability-machine.tex
@@
   \and
-  \inferrule{ }
-  { \revoked \futurestr (\temp,s,\phi_\pub,\phi,H) }
+  \inferrule{ r \in \Regions }
+  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
   \and
@@
 \begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-priv}]
-  % TODO
+  $\readCond{}(\glob)(W)$ picks a $\perma$ region from $W$. $\perma$ regions are persistent over $\futurestr$, so we can use the region that the assumption gives us.
 \end{proof}
@@
   \[
-    \npair{\(start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
+    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
   \]
@@
   \label{lem:writecond-dc}
+  If 
   \begin{itemize}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
-  % TODO
+  $\writeCond{}(\iota,\glob)(W)$ picks a $\perma$ region from $W$. $\perma$ regions are persistent over $\futurestr$, so we can use the region that the assumption gives us.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
-  % TODO
+  Assume $W_2 \futurestr W_1$, $\perm \in \{\exec, \rwx \}$ and $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W_1)$.
+Now let $W_3 \futurestr W_2$, $\addr \in [\start,\addrend]$, and $n' < n$, and show
+  \[
+    \npair{\stdcap[(\perm,\glob)]} \in \stder(W_3)
+  \]
+  by transitivity we have $W_3 \futurestr W_1$, so the result follows from $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W_1)$.
 \end{proof}
@@
   \label{lem:execcond-dc}
+  If
   \begin{itemize}
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
-  % TODO
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-priv}]
+  Assume $W_2 \futurestr W_1$ and $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W_1)$. Now let $W_3 \futurestr W_2$, $n' < n$, and show
+  \[
+    \npair{\stdcap[(\exec,\glob)]} \in \stder(W_3)
+  \]
+  by transitivity we have $W_3 \futurestr W_1$, so the result follows from $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W_1)$.
 \end{proof}
@@
   \label{lem:entrycond-dc}
+  If
   \begin{itemize}
@@
 \begin{lemma}
+  \label{lem:wk-nequal-and-future}
   \begin{align*}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:wk-nequal-and-future}]
+  Construct $W_2'$ as follows:
+  \[
+    W_2(r) =
+    \begin{cases}
+      \revoked & \text{ if $r \in \dom(W_2)$ and $W_1'(r) = \revoked$} \\
+      (v_1',s_1',\phi_{\pub2},\phi_2,H_2) & \text{ if $r \in \dom(W_2)$ and $W_1'(r) = (v_1',s_1',\_,\_,\_)$ and $W_2(r) = (\_,\_,\phi_{\pub2},\phi_2,H_2)$}\\
+      W_1'(r) & \text{ otherwise}
+    \end{cases}
+  \]
+  Notice $\dom(W_2') = \dom(W_1')$.
+\end{proof}
 
 \begin{lemma}
+  \label{lem:str-nequal-and-future}
   \begin{align*}

### 2017-02-01 local-capability-machine.tex
added the W2' worlds needed to prove nequality and future world implies nequal future world

@@
     \begin{cases}
-      \revoked & \text{ if $r \in \dom(W_2)$ and $W_1'(r) = \revoked$} \\
-      (v_1',s_1',\phi_{\pub2},\phi_2,H_2) & \text{ if $r \in \dom(W_2)$ and $W_1'(r) = (v_1',s_1',\_,\_,\_)$ and $W_2(r) = (\_,\_,\phi_{\pub2},\phi_2,H_2)$}\\
+      (v_1',s_1',\phi_{\pub2},\phi_2,H_2) & \arraycolsep=0pt
+      \begin{array}[t]{l}
+        \text{ if $r \in \dom(W_2)$ and $W_1'(r) = (v_1',s_1',\_,\_,\_)$}\\
+        \quad \text{ and $W_2(r) = (\_,\_,\phi_{\pub2},\phi_2,H_2)$}\\
+      \end{array} \\
       W_1'(r) & \text{ otherwise}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:str-nequal-and-future}]
+  Construct $W_2'$ as follows:
+  \[
+    W_2(r) =
+    \begin{cases}
+      (v_1',s_1',\phi_{\pub2},\phi_2,H_2) & \arraycolsep=0pt
+      \begin{array}[t]{l}
+        \text{ if $r \in \dom(W_2)$ and $W_1'(r) = (v_1',s_1',\_,\_,\_)$}\\
+        \quad \text{ and $W_2(r) = (\_,\_,\phi_{\pub2},\phi_2,H_2)$}\\
+      \end{array} \\
+      W_1'(r) & \text{ otherwise}
+    \end{cases}
+  \]
+
+\end{proof}
 

### 2017-02-01 local-capability-machine.tex
Attempt to prove result about sufficiency of the conditions of the load instruction.

--- 2017-01-31 local-capability-machine.tex
@@
   \item $\heapSat[\memheap]{n}{W}$
-  \item $\var{c} = ((\perm',\gl'),\start',\addrend',\addr')$
+  \item $\var{c} = ((\perm,\gl),\start,\addrend,\addr)$
   \item $\npair{c}\in\stdvr(W)$
-  \item $\readAllowed{\perm'}$ 
+  \item $\readAllowed{\perm}$ 
   \item $\withinBounds{\var{c}}$
@@
 
-  then $\npair{\memheap(\addr')} \in \stdvr(W)$
+  then $\npair[n-1]{\memheap(\addr)} \in \stdvr(W)$
 \end{lemma}
 \begin{proof}
-  TODO
+  \begin{enumproof}
+  \item $\npair{(\start,\addrend)} \in \readCond{\gl}(W)$: follows by
+    definition of $\stdvr{\cdot}$ from $\npair{c}\in\stdvr(W)$.
+  \item $\exists r \in \var{localityReg}(g,W)$, $[\start',\addrend'] \supseteq
+    [\start,\addrend] \ldotp$ $W(r)\nsubsim[n-1]
+    \iota_{\start',\addrend'}^\pwl$. By definition of $\readCond{\gl}(W)$. \label{step:unfold-readCond}
+  \item $\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp
+    \memSatPar{\memheap}{W}{P}$. By definition of $\heapSat[\memheap]{n}{W}$.
+  \item $\memheap = \biguplus_{r\in\activeReg{W}}P(r)$ and $\forall r \in
+    \activeReg{W}$, $n'<n \ldotp$ $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
+    and $\npair[n']{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
+    $\memSatPar{\memheap}{W}{P}$. \label{step:unfold-phi-sat-part}
+  \item $r \in \var{localityReg}(g,W)\subseteq \activeReg{W}$. By definition of
+    $\var{localityReg}(\cdot)$ and $\activeReg{\cdot}$.
+  \item $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
+    and $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$.
+    By specializing the result from
+    Step~\ref{step:unfold-phi-sat-part} to the $r$ from
+    Step~\ref{step:unfold-readCond} and $n' = n-1$.
+  \item $\npair[n-1]{P(r)} \in H^\pwl_{\start',\addrend'}(s)(\xi^{-1}(W))$.
+    Follows by combining $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$ with
+    $W(r)\nsubsim[n-1] \iota_{\start',\addrend'}^\pwl$ from Step~\ref{step:unfold-readCond}.
+  \item $\dom(P(r)) = [\start',\addrend']$ and for all $\addr' \in
+    [\start',\addrend']\ldotp$
+    $\npair[n-2]{P(r)(\addr')}\in\stdvr(\xi(\xi^{-1}(W)))$.
+  \item $\addr \in [\start,\addrend]\subseteq[\start',\addrend']$. By combining
+    $\withinBounds(c)$ with the fact that $[\start',\addrend'] \supseteq
+    [\start,\addrend] \ldotp$ from Step~\ref{step:unfold-readCond}.
+  \item In particular, we get: $\memheap(\addr) = P(r)(\addr)$ and
+    $\npair[n-2]{P(r)(\addr)}\in\stdvr(W)$.
+  \end{enumproof}
 \end{proof}

### 2017-02-01 local-capability-machine.tex
Some progress on the FTLR proof.

@@
 
-%make latex preview-mode work with natbib...
-\usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
+% %make latex preview-mode work with natbib...
+% \usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
 
@@
     \end{enumproof}
-  \item Case $\var{conf'} = \Phi'$, and additionally, one of the following
+  \item Case $\var{conf'} = \updateReg[\Phi']{\pcreg}{\var{newPc}}$, and additionally, one of the following
     holds:
@@
     \begin{itemize}
-    \item $\memreg[\Phi'](\pcreg) = \updatePcPerm{\memreg(\lv)}$
-    \item $\memreg[\Phi'](\pcreg) = \var{newPc}$ and $\var{newPc} =
-      ((\perm,\gl),\start,\addrend,\addr + 1)$
-      \dominique{This is wrong: the semantics allow, for example, a move into
-        the PC, after which the new pc is increased by one}
-    \end{itemize}
-    and finally, for all $r \in (\RegName \setminus \{\pcreg\})$, one of the following
+    \item $\var{newPc}= \updatePcPerm{\memreg(\lv)}$
+    \item $\var{newPc} = ((\perm',\gl'),\start',\addrend',\addr' + 1)$ and
+      $\memreg[\Phi'](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$
+    \end{itemize}
+    and finally, for all $r \in \RegName$, one of the following
     holds:
@@
       \end{enumproof}
-    \item $\npair[n-1]{\memreg[\Phi']}\in\stdrr(W)$.
-      \begin{enumproof}
-      \item Assume: $r \in \RegName\setminus\{\pcreg\}$.\\
-        Suffices: $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$.\\
-        By: definition of $\stdrr{\cdot}$
+    \item For all $r \in \RegName$, $\npair[n-1]{\memreg[\Phi'](r)}\in\stdvr(W)$.
+      \begin{enumproof}
       \item Case $\memreg[\Phi'](r) = \memreg[\Phi](r)$:
         $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows from
-        $\npair{\memreg}\in\stdrr(W)$ by definition of $\stdrr(\cdot)$ using
-        Lemma~\ref{lem:stdvr-dc}.
+        Step~\ref{step:ftlr-reg-rr} using Lemma~\ref{lem:stdvr-dc}.
       \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$.
@@
         $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
-        Lemmas~\ref{lem:conds-load-suffice} and~\ref{lem:stdvr-dc} using the
-        fact that $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$
-        which we have from step~\ref{step:ftlr-reg-rr}.
+        Lemmas~\ref{lem:conds-load-suffice} using the fact that
+        $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$ which
+        we have from step~\ref{step:ftlr-reg-rr}.
       \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
@@
       \end{enumproof}
-    \item $\npair[n-1]{\memreg[\Phi'](\pcreg)} \in \stder(W)$: TODO
+    \item $\npair[n-1]{\memreg[\Phi']}\in\stdrr(W)$: Follows from the previous
+      point by definition of $\stdrr(W)$.
+    \item $\npair[n-1]{\var{newPc}} \in \stder(W)$:
+      \begin{enumproof}
+      \item Case $\var{newPc}= \updatePcPerm{\memreg(\lv)}$: We distinguish the
+        following cases:
+        \begin{enumproof}
+        \item Case $\memreg(\lv) = ((\entry,\gl'),\start',\addrend',\addr')$:
+          \begin{enumproof}
+          \item $\npair[n-1]{\memreg(\lv)} \in \stdvr(W)$.  Follows from
+            Step~\ref{step:ftlr-reg-rr} using Lemma~\ref{lem:stdvr-dc}.
+          \item $\npair[n-1]{(\start',\addrend',addr') \in
+              \entryCond{\gl'}(W)}$. By definition of $\stdvr(W)$ from the
+            previous point.
+          \item 
+          \end{enumproof}
+        \end{enumproof}
+      \item Case $\var{newPc} = ((\perm',\gl'),\start',\addrend',\addr' + 1)$ and
+        $\memreg[\Phi'](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$: TODO
+      \end{enumproof}
     \end{enumproof}

### 2017-02-01 local-capability-machine.tex
Added proof of value predicate mono wrt private future for global caps.

--- 2017-02-01 local-capability-machine.tex
@@
 \begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
-  % TODO
+  $\writeCond{}(\iota,\glob)(W)$ picks a $\perma$ region from $W$. $\perma$ regions are persistent over $\futurestr$, so we can use the region that the assumption gives us.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
-  $\writeCond{}(\iota,\glob)(W)$ picks a $\perma$ region from $W$. $\perma$ regions are persistent over $\futurestr$, so we can use the region that the assumption gives us.
+  % TODO
 \end{proof}
@@
   \begin{enumproof}
+  \item $\perm \not\in \{\rwl, \rwlx\}$
   \item $W' \futurestr W$
@@
   \]
-  to this end consider all the possible cases of $\perm$:
+  to this end consider the possible cases of $\perm$ and show that each of the necesary conditions hold:
   \begin{enumproof}
@@
     \item $\perm = \readonly$ \\
-
+      Follows from Lemma~\ref{lem:readcond-mono-priv}.
     \item $\perm = \readwrite$ \\
-
-    \item $\perm = \rwl$ \\
-
+      Follows from Lemma~\ref{lem:readcond-mono-priv} and Lemma~\ref{lem:writecond-mono-priv}.
     \item $\perm = \exec$ \\
-
+      Follows from Lemma~\ref{lem:readcond-mono-priv} and Lemma~\ref{lem:execcond-mono-priv}.
     \item $\perm = \rwx$ \\
-
-    \item $\perm = \rwlx$ \\
-  \end{enumproof}
-
+      Follows from Lemma~\ref{lem:readcond-mono-priv}, Lemma~\ref{lem:writecond-mono-priv}, and Lemma~\ref{lem:execcond-mono-priv}.
+    \item $\perm = \entry$ \\
+      Lemma~\ref{lem:entrycond-mono-priv}
+  \end{enumproof}
 \end{proof}

### 2017-02-01 local-capability-machine.tex

--- 2017-02-01 local-capability-machine.tex
@@
 
-\subsubsection{Downward closedness lemmas}
-\label{sec:downward-closedness}
-
-\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
-  \label{lem:readcond-dc}
-  If
-  \begin{itemize}
-  \item $ n' \leq n$
-  \item $\npair[n]{(\start,\addrend)} \in \readCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair[n']{(\start,\addrend)} \in \readCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
-  \label{lem:writecond-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
-  \end{itemize}
-  then 
-  \[
-    \npair[n']{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{execCondition}$ downwards-closed]
-  \label{lem:execcond-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\perm \in \{\exec, \rwx, \rwlx\}$
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
-  \end{itemize}
-  then 
-  \[
-    \npair[n']{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{enterCondition}$ downwards-closed]
-  \label{lem:entrycond-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
-  \end{itemize}
-  then 
-  \[
-    \npair[n']{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:entrycond-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\stdvr{\cdot}$ downwards-closed]
-  \label{lem:stdvr-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\npair{\word} \in \stdvr(W)$
-  \end{itemize}
-  then 
-  \[
-   \npair[n']{\word} \in \stdvr(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:stdvr-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\stdrr{\cdot}$ downwards-closed]
-  \label{lem:stdrr-dc}
-  \begin{itemize}
-  \item $n' \leq n$
-  \item $\npair{\reg} \in \stdrr(W)$
-  \end{itemize}
-  then 
-  \[
-   \npair[n']{\reg} \in \stdrr(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:stdrr-dc}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[Heap satisfaction downwards closure]
-  \label{lem:heap-sat-dc}
-  \begin{align*}
-    &\forall \hs, n' \leq n, W \ldotp \\
-    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
-  \end{align*}
-\end{lemma}
-
-\subsubsection{Monotonicity lemmas}
-\label{sec:mono}
-
-\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:readcond-mono-pub}
-  If
-  \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{readCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:readcond-mono-priv}
-  If
-  \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\npair{(\start,\addrend)} \in \readCond{}(\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\glob)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-priv}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:writecond-mono-pub}
-  If
-  \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:writecond-mono-priv}
-  If
-  \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
-  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
-  % TODO
-\end{proof}
-
-%ExecCond
-\begin{lemma}[$\mathit{execCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:execcond-mono-pub}
-  If
-  \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\perm \in \{\exec, \rwx, \rwlx\}$
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\perm,\start,\addrend)}\in \execCond{}(\iota,\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{execCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:execcond-mono-priv}
-  If
-  \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\perm \in \{\exec, \rwx \}$
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\perm,\start,\addrend)}\in \execCond{}(\glob)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
-  % TODO
-\end{proof}
-
-% EnterCond
-\begin{lemma}[$\mathit{enterCondition}$ monotone w.r.t $\futurewk$]
-  \label{lem:entrycond-mono-pub}
-  If
-  \begin{itemize}
-  \item $W' \futurewk W$
-  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\iota,\gl)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-pub}]
-  % TODO
-\end{proof}
-
-\begin{lemma}[$\mathit{enterCondition}$ global monotonicity w.r.t $\futurestr$]
-  \label{lem:entrycond-mono-priv}
-  If
-  \begin{itemize}
-  \item $W' \futurestr W$
-  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W)$
-  \end{itemize}
-  then
-  \[
-    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\glob)(W')
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
-  % TODO
-\end{proof}
-
 \subsubsection{Antireduction for the observation relation}
@@
 
+\begin{lemma}[$\mathit{readCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:readcond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:readcond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\npair{(\start,\addrend)} \in \readCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-mono-priv}]
+  $\readCond{}(\glob)(W)$ picks a $\perma$ region from $W$. $\perma$ regions are persistent over $\futurestr$, so we can use the region that the assumption gives us.
+\end{proof}
+
+\begin{lemma}[$\mathit{readCondition}$ downwards-closed]
+  \label{lem:readcond-dc}
+  If
+  \begin{itemize}
+  \item $ n' \leq n$
+  \item $\npair[n]{(\start,\addrend)} \in \readCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair[n']{(\start,\addrend)} \in \readCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:readcond-dc}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:writecond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:writecond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\iota \in \{\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\start,\addrend)}\in \writeCond{}(\iota,\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-mono-priv}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{writeCondition}$ downwards-closed]
+  \label{lem:writecond-dc}
+  If 
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\iota \in \{\iota^\pwl,\iota^\nwl,\iota^{(\nwl,p)}\}$
+  \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:writecond-dc}]
+  $\writeCond{}(\iota,\glob)(W)$ picks a $\perma$ region from $W$. $\perma$ regions are persistent over $\futurestr$, so we can use the region that the assumption gives us.
+\end{proof}
+
+%ExecCond
+\begin{lemma}[$\mathit{execCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:execcond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\perm \in \{\exec, \rwx, \rwlx\}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)}\in \execCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{execCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:execcond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\perm \in \{\exec, \rwx \}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)}\in \execCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-mono-priv}]
+  Assume $W_2 \futurestr W_1$, $\perm \in \{\exec, \rwx \}$ and $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W_1)$.
+Now let $W_3 \futurestr W_2$, $\addr \in [\start,\addrend]$, and $n' < n$, and show
+  \[
+    \npair{\stdcap[(\perm,\glob)]} \in \stder(W_3)
+  \]
+  by transitivity we have $W_3 \futurestr W_1$, so the result follows from $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W_1)$.
+\end{proof}
+
+\begin{lemma}[$\mathit{execCondition}$ downwards-closed]
+  \label{lem:execcond-dc}
+  If
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\perm \in \{\exec, \rwx, \rwlx\}$
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:execcond-dc}]
+  % TODO
+\end{proof}
+
+% EnterCond
+\begin{lemma}[$\mathit{enterCondition}$ monotone w.r.t $\futurewk$]
+  \label{lem:entrycond-mono-pub}
+  If
+  \begin{itemize}
+  \item $W' \futurewk W$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\iota,\gl)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-pub}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[$\mathit{enterCondition}$ global monotonicity w.r.t $\futurestr$]
+  \label{lem:entrycond-mono-priv}
+  If
+  \begin{itemize}
+  \item $W' \futurestr W$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W)$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)}\in \entryCond{}(\glob)(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-priv}]
+  Assume $W_2 \futurestr W_1$ and $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W_1)$. Now let $W_3 \futurestr W_2$, $n' < n$, and show
+  \[
+    \npair{\stdcap[(\exec,\glob)]} \in \stder(W_3)
+  \]
+  by transitivity we have $W_3 \futurestr W_1$, so the result follows from $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\glob)(W_1)$.
+\end{proof}
+
+\begin{lemma}[$\mathit{enterCondition}$ downwards-closed]
+  \label{lem:entrycond-dc}
+  If
+  \begin{itemize}
+  \item $n' \leq n$
+  \item $\npair{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)$
+  \end{itemize}
+  then 
+  \[
+    \npair[n']{(\addr,\start,\addrend)} \in \entryCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:entrycond-dc}]
+  % TODO
+\end{proof}
+
+\subsubsection{LR Sanity lemmas}
+
+\begin{lemma}[Heap satisfaction downwards closure]
+  \label{heap-sat-dc}
+  \begin{align*}
+    &\forall \hs, n' \leq n, W \ldotp \\
+    &\quad \heapSat[\hs]{n}{W} \Rightarrow \heapSat[\hs]{n'}{W}
+  \end{align*}
+\end{lemma}
+
 \subsubsection{Fundamental theorem of logical relations}
@@
   \label{lem:conds-load-suffice}
-  If
-  \begin{itemize}
   \item $\heapSat[\memheap]{n}{W}$
@@
   \item $\withinBounds{\var{c}}$
-  \end{itemize}
-
   then $\npair[n-1]{\memheap(\addr)} \in \stdvr(W)$
@@
    TODO
+ \end{proof}
+
+\begin{lemma}[Fundamental theorem proof helper lemma]
+  \label{lem:ftlr-induction-helper}
+  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
+  If one of the following holds:
+  \begin{itemize}
+  \item \[
+      \begin{gathered}
+        \perm = \exec \land\\
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+      \end{gathered}
+    \]
+  \item \[
+      \begin{gathered}
+        \perm = \rwx \land \\
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
+        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
+      \end{gathered}
+    \]
+  \item \[
+      \begin{gathered}
+        \perm = \rwlx \land\\
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
+        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
+      \end{gathered}
+    \]
+  \end{itemize}
+  and
+  \begin{itemize}
+  \item $\npair[n]{\reg}\in \stdrr(W)$
+  \item $\heapSat[\hs]{n}{W}$
+  \item $i \leq n$
+  \item $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$
+  \end{itemize}
+
+  Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
+  $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
+\end{lemma}
+\begin{proof}
+  \begin{enumproof}
+  \item By induction on $n$. In other words, assume that the theorem already
+    holds for all $n' < n$.
+  \item $i \neq 0$, since
+    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. \\
+    Therefore, assume w.l.o.g. that $i = 1+i'$,
+    \begin{equation*}
+      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step \var{conf'} \step[i'] (\halted,\Phi)
+    \end{equation*}
+  \item $\start \leq \addr \leq \addrend$, $\perm \in \{ \exec,\rwx, \rwlx \}$
+    and $\var{conf'} =
+    \sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+    \uplus \ms_f))$\\
+    By definition of
+    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)
+    \step \var{conf'}$, using the fact that $\var{conf'} \neq \failed$ (since it
+    is not possible that $\failed \step[i'] (\halted,\Phi)$)
+  \item We do a case analysis on
+    $\memheap(a)$
+    \begin{enumproof}
+    \item Case $\memheap(a) = \jmp{\lv}$:
+    \item Case $\memheap(a) = \jnz{\lv}{\rv}$:
+    \item Case $\memheap(a) = \move{\lv}{\rv}$:
+    \item Case $\memheap(a) = \load{\lv}{\hv}$:
+    \item Case $\memheap(a) = \store{\hv}{\rv}$:
+    \item Case $\memheap(a) = \plus{\lv}{\rv}{\rv}$:
+    \item Case $\memheap(a) = \lea{\lv}{\rv}$:
+    \item Case $\memheap(a) = \restricttwo{\lv}{\rv}$:
+    \item Case $\memheap(a) = \subseg{\lv}{\rv}{\rv}$:
+    \item Case $\memheap(a) = \isptr{\lv}{\rv}$:
+    \item Case $\memheap(a) = \getp{\lv}{\lv}$:
+    \item Case $\memheap(a) = \getl{\lv}{\lv}$:
+    \item Case $\memheap(a) = \getb{\lv}{\lv}$:
+    \item Case $\memheap(a) = \gete{\lv}{\lv}$:
+    \item Case $\memheap(a) = \geta{\lv}{\lv}$:
+    \item Case $\memheap(a) = \fail$: Contradiction, since then
+      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+      \uplus \ms_f)) = \failed$, so we have that $\failed \step[i']
+      (\halted,\Phi)$ which is not possible.
+    \item Case $\memheap(a) = \halt$:
+      \begin{enumproof}
+      \item
+        $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+        \uplus \ms_f)) = (\halted, (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
+        \uplus \ms_f))$\\
+        By definition of $\sem{\cdot}$
+      \item Then $i' = 0$ and $\Phi =
+        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+        \ms_f)$. Follows from $(\halted,
+        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
+        \ms_f)) \step[i'] (\halted,\Phi)$
+      \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
+        $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and
+        $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{heap-sat-dc}).
+      \end{enumproof}
+    \end{enumproof}
+  \end{enumproof}
 \end{proof}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:wk-nequal-and-future}]
+  Construct $W_2'$ as follows:
+  \[
+    W_2(r) =
+    \begin{cases}
+      (v_1',s_1',\phi_{\pub2},\phi_2,H_2) & \arraycolsep=0pt
+      \begin{array}[t]{l}
+        \text{ if $r \in \dom(W_2)$ and $W_1'(r) = (v_1',s_1',\_,\_,\_)$}\\
+        \quad \text{ and $W_2(r) = (\_,\_,\phi_{\pub2},\phi_2,H_2)$}\\
+      \end{array} \\
+      W_1'(r) & \text{ otherwise}
+    \end{cases}
+  \]
+  Notice $\dom(W_2') = \dom(W_1')$.
+\end{proof}
 
 \begin{lemma}
+  \label{lem:str-nequal-and-future}
   \begin{align*}

### 2017-02-01 local-capability-machine.tex
global cap vr
Change malloc spec slightly so that c_malloc is hopefully in the value relation.  Specify the latter fact as an (unproven so far) lemma.

@@
                   \begin{aligned}
+                    &c_\malloc = ((\entry,\_),\_,\_,\_) \land\\
                     &\exists \iota_{\malloc,0} \ldotp \\
@@
                     &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
-                    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
+                    &\qquad \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
                     &\qquad \quad \Rightarrow \\
@@
 
+\subsubsection{Malloc safe to pass to adversary}
+\label{sec:malloc-valrel}
+
+\begin{lemma}[Malloc is safe to pass to adversary]
+  If $W(r) = \iota$ with $\iota \futw \iota_{\malloc,0}$, then
+  $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
 \subsubsection{Fundamental theorem of logical relations}

### 2017-02-01 local-capability-machine.tex
Fix typo.
Fix another typo.

@@
 \begin{lemma}[Malloc is safe to pass to adversary]
-  If $W(r) = \iota$ with $\iota \futw \iota_{\malloc,0}$, then
+  If $W(r) = \iota$ with $\iota \futurewk \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
@@
   \item $\addr \in [\start,\addrend]\subseteq[\start',\addrend']$. By combining
-    $\withinBounds(c)$ with the fact that $[\start',\addrend'] \supseteq
+    $\withinBounds{c}$ with the fact that $[\start',\addrend'] \supseteq
     [\start,\addrend] \ldotp$ from Step~\ref{step:unfold-readCond}.

### 2017-02-01 local-capability-machine.tex
Use the right future world relation in the malloc lemma.

@@
 \begin{lemma}[Malloc is safe to pass to adversary]
-  If $W(r) = \iota$ with $\iota \futurewk \iota_{\malloc,0}$, then
+  If $W(r) = \iota$ with $\iota \futurestr \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.

### 2017-02-01 local-capability-machine.tex
Add a TODO about the step we lose too much.
Reremove something that I had deleted, but seems to have found its way back during some merge.

@@
     $\npair[n-2]{P(r)(\addr)}\in\stdvr(W)$.
+    \dominique{TODO: Make sure we lose only one step in the definitions!}
   \end{enumproof}
@@
    TODO
- \end{proof}
-
-\begin{lemma}[Fundamental theorem proof helper lemma]
-  \label{lem:ftlr-induction-helper}
-  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
-  If one of the following holds:
-  \begin{itemize}
-  \item \[
-      \begin{gathered}
-        \perm = \exec \land\\
-        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-      \end{gathered}
-    \]
-  \item \[
-      \begin{gathered}
-        \perm = \rwx \land \\
-        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
-        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
-      \end{gathered}
-    \]
-  \item \[
-      \begin{gathered}
-        \perm = \rwlx \land\\
-        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land\\
-        \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
-      \end{gathered}
-    \]
-  \end{itemize}
-  and
-  \begin{itemize}
-  \item $\npair[n]{\reg}\in \stdrr(W)$
-  \item $\heapSat[\hs]{n}{W}$
-  \item $i \leq n$
-  \item $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$
-  \end{itemize}
-
-  Then $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-  $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n-i}{W'}$
-\end{lemma}
-\begin{proof}
-  \begin{enumproof}
-  \item By induction on $n$. In other words, assume that the theorem already
-    holds for all $n' < n$.
-  \item $i \neq 0$, since
-    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-    \ms_f)\neq (\halted,\Phi)$ for any $\Phi$. \\
-    Therefore, assume w.l.o.g. that $i = 1+i'$,
-    \begin{equation*}
-      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step \var{conf'} \step[i'] (\halted,\Phi)
-    \end{equation*}
-  \item $\start \leq \addr \leq \addrend$, $\perm \in \{ \exec,\rwx, \rwlx \}$
-    and $\var{conf'} =
-    \sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-    \uplus \ms_f))$\\
-    By definition of
-    $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)
-    \step \var{conf'}$, using the fact that $\var{conf'} \neq \failed$ (since it
-    is not possible that $\failed \step[i'] (\halted,\Phi)$)
-  \item We do a case analysis on
-    $\memheap(a)$
-    \begin{enumproof}
-    \item Case $\memheap(a) = \jmp{\lv}$:
-    \item Case $\memheap(a) = \jnz{\lv}{\rv}$:
-    \item Case $\memheap(a) = \move{\lv}{\rv}$:
-    \item Case $\memheap(a) = \load{\lv}{\hv}$:
-    \item Case $\memheap(a) = \store{\hv}{\rv}$:
-    \item Case $\memheap(a) = \plus{\lv}{\rv}{\rv}$:
-    \item Case $\memheap(a) = \lea{\lv}{\rv}$:
-    \item Case $\memheap(a) = \restricttwo{\lv}{\rv}$:
-    \item Case $\memheap(a) = \subseg{\lv}{\rv}{\rv}$:
-    \item Case $\memheap(a) = \isptr{\lv}{\rv}$:
-    \item Case $\memheap(a) = \getp{\lv}{\lv}$:
-    \item Case $\memheap(a) = \getl{\lv}{\lv}$:
-    \item Case $\memheap(a) = \getb{\lv}{\lv}$:
-    \item Case $\memheap(a) = \gete{\lv}{\lv}$:
-    \item Case $\memheap(a) = \geta{\lv}{\lv}$:
-    \item Case $\memheap(a) = \fail$: Contradiction, since then
-      $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-      \uplus \ms_f)) = \failed$, so we have that $\failed \step[i']
-      (\halted,\Phi)$ which is not possible.
-    \item Case $\memheap(a) = \halt$:
-      \begin{enumproof}
-      \item
-        $\sem{\decode(\memheap(\addr))}((\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-        \uplus \ms_f)) = (\halted, (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs
-        \uplus \ms_f))$\\
-        By definition of $\sem{\cdot}$
-      \item Then $i' = 0$ and $\Phi =
-        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-        \ms_f)$. Follows from $(\halted,
-        (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-        \ms_f)) \step[i'] (\halted,\Phi)$
-      \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
-        $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and
-        $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{heap-sat-dc}).
-      \end{enumproof}
-    \end{enumproof}
-  \end{enumproof}
  \end{proof}

### 2017-02-01 local-capability-machine.tex

@@
 \begin{lemma}
+  \label{lem:cap-in-vr-revoketemp-global}
   If
-  \[
-    \npair{\stdcap} \in \stdvr(\revokeTemp{W})
-  \]
+  \begin{itemize}
+  \item $\perm \neq \noperm$
+  \item $\npair{\stdcap} \in \stdvr(\revokeTemp{W})$
+  \end{itemize}
   then 
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:cap-in-vr-revoketemp-global}]
+  Assume for contradiction $\gl = \local$. Procced by case on $\perm$:
+  \begin{enumproof}
+    \item $\perm = \noperm$\\
+      By assumption $\perm \neq \noperm$.
+    \item $\perm \in \{ \readonly, \readwrite, \exec, \rwx, \rwl, \rwlx \}$ \\
+      Here 
+      \[
+        \npair{(\start,\addrend)} \in \readCond{}(local)(\revokeTemp{W})
+      \]
+      must hold, but there are no $\temp$ regions and 
+      \[
+        r \in \localityReg(\local,\revokeTemp)
+      \]
+      must be the region name of a $\temp$ region, so no such $r$ exists. (contradiction)
+    \item $\perm = \entry$ \\
+      % TODO!!! I am not sure it is possible to show for this case.
+    \end{enumproof}
+\end{proof}
+
 \begin{lemma}

### 2017-02-01 local-capability-machine.tex
Changed pub future world to be able to go from revoked to temp region
proof of standard region essential lemmas

@@
     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-  {  (v,s',\phi_\pub',\phi',H') \futurestr (v',s,\phi_\pub,\phi,H) }
+  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
   \and
@@
     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-  {  (v,s',\phi_\pub',\phi',H') \futurewk (v',s,\phi_\pub,\phi,H) }
+  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
+  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
   \and
@@
 \begin{lemma}
-  \label{lem:cap-in-vr-revoketemp-global}
+  \label{lem:revoketemp-readcond}
   If
-  \begin{itemize}
-  \item $\perm \neq \noperm$
-  \item $\npair{\stdcap} \in \stdvr(\revokeTemp{W})$
-  \end{itemize}
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(\revokeTemp{W})
+  \]
   then
   \[
-    \gl = \glob
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
   \]
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:cap-in-vr-revoketemp-global}]
-  Assume for contradiction $\gl = \local$. Procced by case on $\perm$:
-  \begin{enumproof}
-    \item $\perm = \noperm$\\
-      By assumption $\perm \neq \noperm$.
-    \item $\perm \in \{ \readonly, \readwrite, \exec, \rwx, \rwl, \rwlx \}$ \\
-      Here 
-      \[
-        \npair{(\start,\addrend)} \in \readCond{}(local)(\revokeTemp{W})
-      \]
-      must hold, but there are no $\temp$ regions and 
-      \[
-        r \in \localityReg(\local,\revokeTemp)
-      \]
-      must be the region name of a $\temp$ region, so no such $r$ exists. (contradiction)
-    \item $\perm = \entry$ \\
-      % TODO!!! I am not sure it is possible to show for this case.
-    \end{enumproof}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-readcond}]
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(\revokeTemp{W})
+  \]
+  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
+  \[
+    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n-1] \iota^\pwl_{[\start',\addrend']}
+  \]
+  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
 \end{proof}
@@
 \begin{lemma}
+  \label{lem:revoketemp-writecond}
   If
   \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
+  \]
+  then
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
+  \]
+  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
+  \[
+    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
+  \]
+  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-execcond}
+  If
+  \begin{itemize}
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(\revokeTemp{W})$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-execcond}]
+  Use Lemma~\ref{lem:rt-w-pub-future-w}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-entercond}
+  If
+  \begin{itemize}
+  \item $\npair{(\addr,\start,\addrend)} \in \execCond{}(\gl)(\revokeTemp{W})$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-entercond}]
+    Use Lemma~\ref{lem:rt-w-pub-future-w}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-vr-implies-vr}
+  If
+  \[
     \npair{w} \in \stdvr(\revokeTemp{W})
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-vr-implies-vr}]
+  Follows from Lemma~\ref{lem:revoketemp-readcond}, Lemma~\ref{lem:revoketemp-writecond}, Lemma~\ref{lem:revoketemp-execcond}, and Lemma~\ref{lem:revoketemp-entercond}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:hwnl-nsubset-hpwl}
+  \[
+    H^\nwl_{\start,\addrend} \; s\; \hat{W} \nsubseteq H^\pwl_{\start,\addrend} \; s\; \hat{W}
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:hwnl-nsubset-hpwl}]
+  Let
+  \[
+    \npair{\ms} \in H^\nwl_{\start,\addrend} \; s\; \hat{W}
+  \]
+  and show
+  \[
+    \npair{\ms} \in H^\pwl_{\start,\addrend} \; s\; \hat{W}
+  \]
+  From the assumption, we get $\dom(\ms) = [\start,\addrend]$. We further need to show
+  \[
+    \forall \addr \in \dom(\ms) \ldotp \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))
+  \]
+  Given $\addr$, we know from the assumption that
+  \[
+    \npair[n-1]{\ms(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
+  \]
+  Now the desired reulst follows from Lemma~\ref{lem:revoketemp-vr-implies-vr}.
+\end{proof}
 
@@
   \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp  \\
+    & \forall n \in \nats\ldotp \forall \start, \addrend \in \Addrs \ldotp  \\
     & \quad \iota^\nwl_{\start,\addrend} \nsubsim \iota^\pwl_{\start,\addrend}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:nwl-subset-pwl}]
-
+  Let $n$, $\start$, $\addrend$ be given and show
+  \[
+    \iota^\nwl_{\start,\addrend} \nsubsim \iota^\pwl_{\start,\addrend}
+  \]
+  They agree on the state and transition systems, so given $\hat{W}$ it suffices to show
+  \[
+    H^\nwl_{\start,\addrend} \; 1\; \hat{W} \nsubseteq H^\pwl_{\start,\addrend} \; 1\; \hat{W}  
+  \]
+  which is true by Lemma~\ref{lem:hwnl-nsubset-hpwl}.
 \end{proof}
@@
   \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp  \\
+    & \forall n \in \nats\ldotp \forall \start, \addrend \in \Addrs \ldotp  \\
     & \quad \iota^{\nwl,p}_{\start,\addrend}  \nsubsim \iota^\pwl_{\start,\addrend}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:nwlp-subset-pwl}]
-
+  Follows from Lemma~\ref{lem:hwnl-nsubset-hpwl}.
 \end{proof}
@@
   \begin{align*}
-    & \forall n \in \nats\ldotp \forall \addr, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
-    & \quad \dom(\ms) = [\star,\addrend] \Rightarrow \\
+    & \forall n \in \nats\ldotp \forall \start, \addrend \in \Addrs \ldotp \forall v \in \{\perma, \temp\} \ldotp \\
+    & \quad \dom(\ms) = [\start,\addrend] \Rightarrow \\
     & \qquad \iota^{\sta,u}_{\start,\addrend}(v,\ms) \nsubsim \iota^\pwl_{\start,\addrend}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:stau-subset-pwl}]
-
+  Essentially the same as the proof of Lemma~\ref{lem:nwl-subset-pwl} and Lemma~\ref{lem:hwnl-nsubset-hpwl}.
 \end{proof}
@@
 \subsubsection{Future worlds}
+\begin{lemma}[World public future world of revoked world]
+  \label{lem:rt-w-pub-future-w}
+  \[
+    \forall W \ldotp \revokeTemp{W} \futurewk W
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:rt-w-pub-future-w}]
+  For all $r$ where $W(r) = (\temp,s,\phi_\pub,\phi,H)$, we have $\revokeTemp{W} = \revoked$. By the public future region relation we have
+  \[
+    W(r) = (\temp,s,\phi_\pub,\phi,H) \futurewk \revokeTemp{W}(r) = \revoked
+  \]
+  all other regions remain unchanged, so this follows by reflexivity of the public future region relation.
+\end{proof}
+
 \begin{lemma}

### 2017-02-01 local-capability-machine.tex
Added missing TODOs

@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:priv-mono-like}]
+  % TODO
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:priv-mono-like2}]
+  % TODO
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}[lem:revoke-temp-stack]
+  % TODO
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:disj-mem-sat}]
+  % TODO
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:mem-sat-static}]
+  % TODO
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:mem-sat-data-only-std-regions}]
+  % TODO
+\end{proof}
 
@@
 
-\begin{lemma}
-  For all $W$, $W'$, and $W''$
-  \begin{itemize}
-  \item $W' \futurewk W \Rightarrow W' \futurestr W$
-  \item $\wrev{W} \futurestr W$
-  \end{itemize}
-  Combined with transitivity of $\futurestr$, the above gives us:
-  \begin{itemize}
-  \item $W'' \futurestr W' \land W' \futurewk W \Rightarrow W'' \futurestr W$
-  \item $W'' \futurewk W' \land W' \futurestr W \Rightarrow W'' \futurestr W$
-  \end{itemize}
-\end{lemma}
+\begin{lemma}[World private future world of revoked world]
+  \label{lem:rt-w-priv-future-w}
+  \[
+    \forall W \ldotp \wrev{W} \futurestr W
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:rt-w-priv-future-w}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[Public future world relation included in private future world relation]
+  \label{lem:future-pub-impl-future-priv}
+  \[
+   W' \futurewk W \Rightarrow W' \futurestr W
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:future-pub-impl-future-priv}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[Transitivity proberties between private and public future worlds]
+\[
+  W'' \futurestr W' \land W' \futurewk W \Rightarrow W'' \futurestr W
+\]
+and
+\[
+  W'' \futurewk W' \land W' \futurestr W \Rightarrow W'' \futurestr W
+\]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:future-priv-pub-trans}]
+  % TODO
+\end{proof}
 

### 2017-02-01 local-capability-machine.tex
Latex fix, add note to RW.
Fix another merge artifact.
Fix the last fix.
Some more work on the FTLR proof.

@@
 
+\begin{lemma}[Failing terms are in $\observations$ and $\stder$]
+  \label{lem:failed-obs-stder}
+  If $(\reg,\ms\uplus\ms_f) \step[*] \failed$ for all $\ms_f$, then
+  $\npair{(\reg,\ms)} \in \observations(W)$ for any $W$.
+
+  If $(\reg\update{\pcreg}{\word},\ms) \step[*] \failed$ for all $\reg, \ms$,
+  then $\npair{\word} \in \stder(W)$ for any $W$.
+\end{lemma}
+\begin{proof}
+  Follows from the definitions of $\observations(W)$ and $\stder(W)$ using an
+  (omitted) determinacy result. \dominique{at least state the result
+    somewhere?}
+\end{proof}
+
 \begin{lemma}[Anti-reduction for $\observations$]
@@
   \label{lem:conds-load-suffice}
+  If
+  \begin{itemize}
   \item $\heapSat[\memheap]{n}{W}$
@@
   \item $\withinBounds{\var{c}}$
+  \end{itemize}
   then $\npair[n-1]{\memheap(\addr)} \in \stdvr(W)$
@@
   \begin{enumproof}
+  \item By induction on $n$. In other words, assume that the theorem already
+    holds for all $n' < n$.
   \item Assume: $n' \leq n$, $\npair[n']{\reg}\in \stdrr(W)$, $\heapSat[\hs]{n'}{W}$.\\
@@
     By: definition of $\observations(W)$ \label{suff-after-eval}
-  \item By induction on $n$. In other words, assume that the theorem already
-    holds for all $n' < n$.
   \item $i \neq 0$, since
@@
       \Phi \step \var{conf'} \step[i'] (\halted,\Phi')
-    \end{equation*}
-  \item $n > 0$, since otherwise $i = 0$ (because $i \leq n$) and this is
+    \end{equation*} \label{step:ip-non-zero}
+  \item $n' > 0$, since otherwise $i = 0$ (because $i \leq n'\leq n$) and this is
     impossible by the previous point.
-  \item $\npair{\memreg(\pcreg)}\in\stdvr(W)$. \label{step:ftlr-pc-vr}
+  \item $\npair[n']{\memreg(\pcreg)}\in\stdvr(W)$. \label{step:ftlr-pc-vr}
     Proof:
@@
       \sqsubseteq\perm$\\
-      Suffices: $\npair{(\perm',\start,\addrend)}\in\execCond{g}(W)$\\
+      Suffices: $\npair[n']{(\perm',\start,\addrend)}\in\execCond{g}(W)$\\
       By: the definition of $\stdvr(\cdot)$ using the assumptions
-    \item Assume: $n'<n$, $W'\future W$, $\addr' \in [\start,\addrend]$, $\gl
+    \item Assume: $n''<n'$, $W'\future W$, $\addr' \in [\start,\addrend]$, $\gl
       = \local \Rightarrow \future = \futurewk$,
       $\gl=\glob\Rightarrow\future = \futurestr$.\\
-      Suffices: $\npair[n']{((\perm,\gl),\start,\addrend,\addr')} \in
+      Suffices: $\npair[n'']{((\perm,\gl),\start,\addrend,\addr')} \in
       \stder(W')$.
@@
     \end{enumproof}
-  \item For all $r \in \RegName$, $\npair[n]{\memreg(r)}\in\stdvr(W)$. \label{step:ftlr-reg-rr}
-    \begin{enumproof}
-    \item Case $r \neq \pcreg$: follows from $\npair[n]{\reg}\in \stdrr(W)$ by
+  \item For all $r \in \RegName$, $\npair[n']{\memreg(r)}\in\stdvr(W)$. \label{step:ftlr-reg-rr}
+    \begin{enumproof}
+    \item Case $r \neq \pcreg$: follows from $\npair[n']{\reg}\in \stdrr(W)$ by
       definition of $\stdrr(W)$.
@@
       $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and
-      $\heapSat[\hs']{n-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
-    \end{enumproof}
-  \item Case $\var{conf'} = \updateReg[\Phi']{\pcreg}{\var{newPc}}$, and additionally, one of the following
+      $\heapSat[\hs']{n'-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
+    \end{enumproof}
+  \item Case $\var{conf'} = \updateReg[\Phi'']{\pcreg}{\var{newPc}}$, and additionally, one of the following
     holds:
     \begin{itemize}
-    \item $\memheap[\Phi'] = \memheap[\Phi]$
-    \item $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$, with
+    \item $\memheap[\Phi''] = \memheap[\Phi]$
+    \item $\memheap[\Phi''] = \memheap[\Phi]\update{\addr'}{\var{w}}$, with
       $\memreg[\Phi](r_1) = ((\perm',\gl'),\start',\addrend',\addr') = c$ and
@@
     \item $\var{newPc} = ((\perm',\gl'),\start',\addrend',\addr' + 1)$ and
-      $\memreg[\Phi'](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$
+      $\memreg[\Phi''](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$
     \end{itemize}
@@
     \begin{itemize}
-    \item $\memreg[\Phi'](r) = \memreg[\Phi](r)$
-    \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$
-    \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =
+    \item $\memreg[\Phi''](r) = \memreg[\Phi](r)$
+    \item $\memreg[\Phi''](r) = z$ for some $z \in \ints$
+    \item $\memreg[\Phi''](r) = \var{w}$ and $\memreg(r_2) =
       ((\perm',\gl'),\start',\addrend',\addr') = \var{c}$ and
@@
       \memheap(\addr')$
-    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
+    \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r_1) =
       ((\perm',\gl'),\start',\addrend',\addr')$ and $\perm' \neq \entry$ and
-      $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in
+      $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + z)$ for some $z \in
       \ints$
-    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+    \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r) =
       ((\perm',\gl'),\start',\addrend',\addr')$ and $(\perm'',\gl'')\sqsubseteq
       (\perm',\gl')$ and $c = ((\perm'',\gl''),\start',\addrend',\addr')$
-    \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+    \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r) =
       ((\perm',\gl'),\start',\addrend',\addr')$ and $\start' \leq \start''$ and
@@
     \begin{enumproof}
-    \item $\heapSat[{\memheap[\Phi']}]{n-1}{W}$.
-      \begin{enumproof}
-      \item Case $\memheap[\Phi'] = \memheap[\Phi]$:
-        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
-      \item Case $\memheap[\Phi'] = \memheap[\Phi]\update{\addr'}{\var{w}}$,
+    \item $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$.
+      \begin{enumproof}
+      \item Case $\memheap[\Phi''] = \memheap[\Phi]$:
+        $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
+      \item Case $\memheap[\Phi''] = \memheap[\Phi]\update{\addr'}{\var{w}}$,
         with $\memreg[\Phi](r_1) = ((\perm',\gl'),\start',\addrend',\addr') = c$
@@
         \in \{\rwlx,\readwritel \}$.\\
-        $\heapSat[{\memheap[\Phi']}]{n-1}{W}$ follows by
+        $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$ follows by
         Lemmas~\ref{lem:conds-store-suff} and~\ref{lem:heap-sat-dc} using the
-        fact that $\heapSat[\memheap]{n}{W}$ and
-        $\npair{\memreg(r_1)}\in\stdvr(W)$ and
-        $\npair{\memreg(r_2)}\in\stdvr(W)$ which follows from
-        $\npair{\memreg}\in\stdrr(W)$ by definition.
-      \end{enumproof}
-    \item For all $r \in \RegName$, $\npair[n-1]{\memreg[\Phi'](r)}\in\stdvr(W)$.
-      \begin{enumproof}
-      \item Case $\memreg[\Phi'](r) = \memreg[\Phi](r)$:
-        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows from
+        fact that $\heapSat[\memheap]{n'}{W}$ and
+        $\npair[n']{\memreg(r_1)}\in\stdvr(W)$ and
+        $\npair[n']{\memreg(r_2)}\in\stdvr(W)$ which follows from
+        $\npair[n']{\memreg}\in\stdrr(W)$ by definition.
+      \end{enumproof}
+    \item For all $r \in \RegName$, $\npair[n'-1]{\memreg[\Phi''](r)}\in\stdvr(W)$.\label{step:phip-regs-stdvr}
+      \begin{enumproof}
+      \item Case $\memreg[\Phi''](r) = \memreg[\Phi](r)$:
+        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows from
         Step~\ref{step:ftlr-reg-rr} using Lemma~\ref{lem:stdvr-dc}.
-      \item $\memreg[\Phi'](r) = n$ for some $n \in \ints$.
-        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by definition of
+      \item $\memreg[\Phi''](r) = z$ for some $z \in \ints$.
+        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by definition of
         $\stdvr(\cdot)$
-      \item $\memreg[\Phi'](r) = \var{w}$ and $\memreg(r_2) =
+      \item $\memreg[\Phi''](r) = \var{w}$ and $\memreg(r_2) =
         ((\perm',\gl'),\start',\addrend',\addr') = \var{c}$ and
@@
         \memheap(\addr')$:\\
-        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by
         Lemmas~\ref{lem:conds-load-suffice} using the fact that
-        $\heapSat[\memheap]{n}{W}$ and $\npair{\memreg(r_2)}\in\stdvr(W)$ which
+        $\heapSat[\memheap]{n'}{W}$ and $\npair[n']{\memreg(r_2)}\in\stdvr(W)$ which
         we have from step~\ref{step:ftlr-reg-rr}.
-      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r_1) =
+      \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r_1) =
         ((\perm',\gl'),\start',\addrend',\addr')$ and $\perm' \neq \entry$ and
-        $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + n)$ for some $n \in
+        $\var{c} = ((\perm',\gl'),\start',\addrend',\addr' + z)$ for some $z \in
         \ints$: \\
-        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by
         Lemmas~\ref{lem:conds-lea-suffice} and~\ref{lem:stdvr-dc} using the
-        fact that $\npair{\memreg(r_1)}\in\stdvr(W)$
+        fact that $\npair[n']{\memreg(r_1)}\in\stdvr(W)$
         which we have from step~\ref{step:ftlr-reg-rr}.
-      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+      \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r) =
         ((\perm',\gl'),\start',\addrend',\addr')$ and $(\perm'',\gl'')\sqsubseteq
         (\perm',\gl')$ and $c = ((\perm'',\gl''),\start',\addrend',\addr')$: \\
-        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by
         Lemmas~\ref{lem:conds-restrict-suffice} and~\ref{lem:stdvr-dc} using the
-        fact that $\npair{\memreg(r)}\in\stdvr(W)$
-        which follows from $\npair{\memreg}\in\stdrr(W)$ by
+        fact that $\npair[n']{\memreg(r)}\in\stdvr(W)$
+        which follows from $\npair[n']{\memreg}\in\stdrr(W)$ by
         definition.
-      \item $\memreg[\Phi'](r) = \var{c}$ and $\memreg(r) =
+      \item $\memreg[\Phi''](r) = \var{c}$ and $\memreg(r) =
         ((\perm',\gl'),\start',\addrend',\addr')$ and $\start' \leq \start''$ and
@@
         ((\perm',\gl'),\start'',\addrend'',\addr')$: \\
-        $\npair[n-1]{\memreg[\Phi'](r)} \in \stdvr(W)$ follows by
+        $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by
         Lemmas~\ref{lem:conds-subseg-suffice} and~\ref{lem:stdvr-dc} using the
-        fact that $\npair{\memreg(r)}\in\stdvr(W)$
-        which follows from $\npair{\memreg}\in\stdrr(W)$ by definition.
-      \end{enumproof}
-    \item $\npair[n-1]{\memreg[\Phi']}\in\stdrr(W)$: Follows from the previous
+        fact that $\npair[n']{\memreg(r)}\in\stdvr(W)$
+        which follows from $\npair[n']{\memreg}\in\stdrr(W)$ by definition.
+      \end{enumproof}
+    \item $\npair[n'-1]{\memreg[\Phi'']}\in\stdrr(W)$: Follows from the previous
       point by definition of $\stdrr(W)$.
-    \item $\npair[n-1]{\var{newPc}} \in \stder(W)$:
+    \item $\npair[n'-1]{\var{newPc}} \in \stder(W)$:
       \begin{enumproof}
@@
           \begin{enumproof}
-          \item $\npair[n-1]{\memreg(\lv)} \in \stdvr(W)$.  Follows from
-            Step~\ref{step:ftlr-reg-rr} using Lemma~\ref{lem:stdvr-dc}.
-          \item $\npair[n-1]{(\start',\addrend',addr') \in
-              \entryCond{\gl'}(W)}$. By definition of $\stdvr(W)$ from the
+          \item $\npair[n']{\memreg(\lv)} \in \stdvr(W)$.  Follows from
+            Step~\ref{step:ftlr-reg-rr}.
+          \item $\npair[n']{(\start',\addrend',addr')} \in \entryCond{\gl'}(W)$.
+            By definition of $\stdvr(W)$ from the previous point.
+          \item $\npair[n'-1]{((\exec,\gl'),\start',\addrend',\addr')} \in
+            \stder(W)$: By definition of $\entryCond{\cdot}$ and taking $n' =
+            n'-1$ and $W' = W$
+          \item $\updatePcPerm{\memreg(\lv)} =
+            ((\exec,\gl'),\start',\addrend',\addr')$: by definition of $\updatePcPerm{\cdot}$.
+          \end{enumproof}
+        \item Case $\memreg(\lv) = ((\perm',\gl'),\start',\addrend',\addr')$
+          with $\perm' \in \{\exec,\rwx,\rwlx\}$ and $\withinBounds{\memreg(\lv)}$:
+          \begin{enumproof}
+          \item $\npair[n']{\memreg(\lv)} \in \stdvr(W)$.  Follows from
+            Step~\ref{step:ftlr-reg-rr}.
+          \item $\npair[n']{(\perm',\start',\addrend',\addr')} \in \execCond{\gl'}(W)$.
+            By definition of $\stdvr(W)$ from the previous point.
+          \item $\npair[n'-1]{((\perm',\gl'),\start',\addrend',\addr')} \in
+            \stder(W)$: By definition of $\execCond{\cdot}$, taking $n' = n'-1$,
+            $W' = W$ and $\addr = \addr'$. Note that $\addr' \in
+            [\start',\addrend']$ because we have $\withinBounds{\memreg(\lv)}$.
+          \item $\updatePcPerm{\memreg(\lv)} =
+            ((\perm',\gl'),\start',\addrend',\addr')$: by definition of
+            $\updatePcPerm{\cdot}$.
+          \end{enumproof}
+        \item Case not ($\memreg(\lv) =
+          ((\entry,\gl'),\start',\addrend',\addr')$) and not ($\memreg(\lv) =
+          ((\perm',\gl'),\start',\addrend',\addr')$ with $\perm' \in
+          \{\exec,\rwx,\rwlx\}$ and $\withinBounds{\memreg(\lv)}$):
+          \begin{enumproof}
+          \item $\updatePcPerm{\memreg(\lv)} = \memreg(\lv)$: by definition of 
+            $\updatePcPerm{\cdot}$.
+          \item $(\reg\update{\pcreg}{\memreg(\lv)},\ms)\step\failed$ for any
+            $\reg$,$\ms$: by definition of the evaluation relation.
+          \item $\npair[n'-1]{\var{newPc}} \in \stder(W)$: by
+            Lemma~\ref{lem:failed-obs-stder} using the previous point.
+          \end{enumproof}
+        \end{enumproof}
+      \item Case $\var{newPc} = ((\perm',\gl'),\start',\addrend',\addr' + 1)$ and
+        $\memreg[\Phi''](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$:
+        \begin{enumproof}
+        \item Case $\perm' \in \{\exec,\rwx,\rwlx\}$ and $\start'\leq
+          \addr'+1\leq \addrend'$:
+          \begin{enumproof}
+          \item $\npair[n'-1]{\memreg[\Phi''](\pcreg)} \in \stdvr(W)$: by
+            Step~\ref{step:phip-regs-stdvr}.
+          \item $\npair[n'-1]{((\perm',\gl'),\start',\addrend',\addr' + 1)} \in \stdvr(W)$: by
+            Lemma~\ref{lem:conds-lea-suffice} from the previous point.
+          \item One of the following holds:
+            \begin{itemize}
+            \item \[
+                \begin{gathered}
+                  \perm' = \exec \land\\
+                  \npair[n'-1]{(\start',\addrend')} \in \readCond{}(\gl)(W)
+                \end{gathered}
+              \]
+            \item \[
+                \begin{gathered}
+                  \perm' = \rwx \land \\
+                  \npair[n'-1]{(\start',\addrend')} \in \readCond{}(\gl)(W) \land\\
+                  \npair[n'-1]{(\start',\addrend')} \in \writeCond{}(\iota^\nwl,\gl)(W)
+                \end{gathered}
+              \]
+            \item \[
+                \begin{gathered}
+                  \perm' = \rwlx \land\\
+                  \npair[n'-1]{(\start',\addrend')} \in \readCond{}(\gl)(W) \land\\
+                  \npair[n'-1]{(\start',\addrend')} \in \writeCond{}(\iota^\pwl,\gl)(W),
+                \end{gathered}
+              \]
+            \end{itemize}
+           This follows from the previous point by  definition of $\stdvr(W)$
+         \item $\npair[n'-1]{((\perm',\gl'),\start',\addrend',\addr' + 1)} \in
+           \stder(W)$: By the induction hypothesis of this lemma using the
            previous point.
+          \end{enumproof}
+        \item Case not ($\perm' \in \{\exec,\rwx,\rwlx\}$ and $\start'\leq
+          \addr'+1\leq \addrend'$):
+          The result follows by Lemma~\ref{lem:failed-obs-stder}.
+        \end{enumproof}
+      \end{enumproof}
     \item
-          \end{enumproof}
-        \end{enumproof}
-      \item Case $\var{newPc} = ((\perm',\gl'),\start',\addrend',\addr' + 1)$ and
-        $\memreg[\Phi'](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$: TODO
-      \end{enumproof}
+      $\npair[n'-1]{(\memreg[\Phi'']\update{\pcreg}{\var{newPc}},\memheap[\Phi''])}\in\observations(W)$,
+      i.e. $\npair[n'-1]{\var{conf'}}\in\observations(W)$:
+      by definition of $\stder(W)$ using the above three points.
+    \item $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
+      $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+      By: definition of $\observations(W)$ using the previous step and the
+      evaluation $\var{conf'}\step[i'] (\halted,\Phi')$ from
+      Step~\ref{step:ip-non-zero}.\dominique{TODO: check.  What about ms_f and other stuff.}
     \end{enumproof}
@@
 language with object capabilities.  Relevant because it treats object
-capabilities, albeit in a JavaScript-like lambda calculus.
+capabilities, albeit in a JavaScript-like lambda calculus. It also deals with an
+untyped language, using a semantic unitype.
 

### 2017-02-02 local-capability-machine.tex
missing $
removed n' < n requirement from memSat
Had forgotten to change a n' i previous commit
Added proof of memsat downwards closure

@@
         &\forall r \in \activeReg{W} \ldotp\\
-        &\quad  \forall n'<n \ldotp \\
-        &\qquad \exists H,s \ldotp\\
-        &\qquad \quad W(r) = (\_,s,\_,\_,H) \land \\
-        &\qquad \quad \npair[n']{P(r)} \in H(s)(\xi^{-1}(W))\\
+        &\quad \exists H,s \ldotp\\
+        &\qquad W(r) = (\_,s,\_,\_,H) \land \\
+        &\qquad \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
       \end{aligned}\right.
@@
 \begin{lemma}[Heap satisfaction downwards closure]
-  \label{heap-sat-dc}
+  \label{lem:heap-sat-dc}
   \begin{align*}
@@
 \end{lemma} 
+\begin{proof}[Proof of Lemma~\ref{lem:heap-sat-dc}]
+  Let $\ms$, $n' \leq n$, and $W$ be given and assume
+  \[
+    \heapSat[\hs]{n}{W}
+  \]
+  This assumption gives us $P : \activeReg{W} \fun \MemSegments$ such that
+  \begin{enumproof}
+    \item $\hs = \biguplus_{r\in\activeReg{W}}P(r)$ \label{pf:heap-sat:prop1}
+    \item \label{pf:heap-sat:prop2}
+      \begin{align*}
+        &\forall r \in \activeReg{W} \ldotp\\
+        &\quad \exists H,s \ldotp\\
+        &\qquad W(r) = (\_,s,\_,\_,H) \land \\
+        &\qquad \npair[n']{P(r)} \in H(s)(\xi^{-1}(W))
+      \end{align*}
+  \end{enumproof}
+  Using $P$ as witness, \ref{pf:heap-sat:prop1} is the first condition we need. Now let $r$ be given and use \ref{pf:heap-sat:prop2} to get $H$ and $s$ such that
+  \begin{enumproof}[resume]
+    \item $W(r) = (\_,s,\_,\_,H)$ 
+    \item $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$ \label{pf:heap-sat:interp}
+  \end{enumproof}
+We now need to show
+\[
+        \npair[n']{P(r)} \in H(s)(\xi^{-1}(W))
+\]
+which follows from \ref{pf:heap-sat:interp}, $n' \leq n$, and $H(s)(\xi^{-1}(W))$ is a $\UPred{\MemSegments}$.
+\end{proof}
 
@@
       evaluation $\var{conf'}\step[i'] (\halted,\Phi')$ from
-      Step~\ref{step:ip-non-zero}.\dominique{TODO: check.  What about ms_f and other stuff.}
+      Step~\ref{step:ip-non-zero}.\dominique{TODO: check.  What about $ms_f$ and other stuff.}
     \end{enumproof}

### 2017-02-02 local-capability-machine.tex
fixed some proofs that were explicit about memsat and added malloc spec assumption to lemma about awkward example

@@
       \item $\ms_\adv$:\\
-        let $n' < n$ be given and show
-        \[
-          \npair[n']{\ms_\adv} \in H^\pwl_{\start_\adv,\addrend_\adv}\;1\;\xi^{-1}(W)
+        show
+        \[
+          \npair[n]{\ms_\adv} \in H^\pwl_{\start_\adv,\addrend_\adv}\;1\;\xi^{-1}(W)
         \]
@@
         \[
-          \npair[n'-1]{c_\link} \in \stdvr(W)
+          \npair[n-1]{c_\link} \in \stdvr(W)
         \]
@@
         \[
-          \npair[n'-1]{\link,\link+1} \in \readCond{}(\glob)(W)
+          \npair[n-1]{\link,\link+1} \in \readCond{}(\glob)(W)
         \]
@@
         \begin{enumproof}
-          \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: \label{item:f3:adv1}\\
+          \item $\npair[n-1]{c_\adv} \in \stdvr(W)$: \label{item:f3:adv1}\\
             We will show
@@
             $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}$ governs the adversary, so the result follows from Lemma~\ref{lem:nwlp-subset-pwl}.
-          \item $\npair[n'-1]{c_\malloc} \in \stdvr(W)$\\
-            Follows from the malloc specification.
+          \item $\npair[n-1]{c_\malloc} \in \stdvr(W)$\\
+            Which follows from Lemma~\ref{lem:malloc-in-vr}.
             \lau{I don't think this is actually the case at the moment. I guess we either need to show this based on the specification or just require it in the specification.}
@@
   where 
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc
+  \end{itemize}
   \begin{align*}
@@
 \begin{lemma}[Malloc is safe to pass to adversary]
+  \label{lem:malloc-in-vr}
   If $W(r) = \iota$ with $\iota \futurestr \iota_{\malloc,0}$, then

### 2017-02-02 local-capability-machine.tex
added more malloc specification assumptions to lemmas
proof of existence theorem

@@
                 \end{align*}
+                and
+                \begin{itemize}
+                \item $c_\malloc$ satisfies the specification for malloc.
+                \end{itemize}
                 where 
@@
   \end{align*}
+  and
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc.
+  \end{itemize}
   where 
@@
 
+\begin{comment}
 \begin{lemma}\forcenewline
@@
 \end{lemma}
+\end{comment}
 
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:world-existence}
   There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are preordered \cofes{} and there exists an isomorphism $\xi$ such that
@@
 \end{mathpar}
+
+% XXX
+
+\begin{proof}[Proof of Theorem~\ref{thm:world-existence}]
+  The theorem follows from a more general solution theorem for the
+  category of $P$ preordered cofe's~\cite{bizjak-note}. 
+  We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to
+  $P$. 
+  \begin{align*}
+     F_1((X,\futurestr'),(Y,\futurewk')) = \\
+     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
+                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,\futurewk') \monnefun \UPred{\HeapSegments})) + \\
+                                          & \{\perma\} \times \States
+                                            \times \Rels \times
+                                            (\States \fun ((X,\futurestr')
+                                            \monnefun
+                                            \UPred{\HeapSegments})))),
+                                        \futurestr)
+  \end{align*}
+  and
+  \begin{align*}
+     F_2((X,\futurestr'),(Y,\futurewk')) = \\
+     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
+                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,\futurewk') \monnefun \UPred{\HeapSegments})) + \\
+                                          & \{\perma\} \times \States
+                                            \times \Rels \times
+                                            (\States \fun ((X,\futurestr')
+                                            \monstrnefun
+                                            \UPred{\HeapSegments})))),
+                                        \futurestr)
+  \end{align*}
+  The orderings $\futurestr$ and $\futurewk$ used in the definition of
+  $F_1$ and $F_2$ are defined by the properties given above.
+  Note that the image of $F_1$ and $F_2$ only differ in the ordering
+  relation, i.e., letting $U$ denote the forgetful functor from the
+  category of preordered cofe's to the category of cofe's, we have
+  $U \circ F_1 = U \circ F_2$.
+  From~\cite{bizjak-note} it then follows that there exists a cofe $\Wor$ and two
+  preorderings $\futurestr$ and $\futurewk$ and an isomorphism $\xi$ satisfying the properties
+  claimed in theorem.
+\end{proof}
+
+% XXX end
 

### 2017-02-02 local-capability-machine.tex
proof of existence theorem
added lemmas about regions

--- 2017-02-02 local-capability-machine.tex
@@
 %
+In the proof of Lemma~\ref{lem:correctness-g1}, we will use the following region
+\begin{definition}
+  \label{def:iotax-region}
+  \begin{align*}
+    \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
+    \phi_\pub & = \{(0,1)\}^* \\
+    \phi      & = (1,0) \union \phi_\pub \\
+    H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \uplus \{\npair[0]{\ms}\}
+  \end{align*}
+\end{definition}
+\begin{lemma}
+  \label{lem:iotax-is-a-region}
+  The region in Definition~\ref{def:iotax-region} is a region.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:iotax-is-a-region}]
+
+\end{proof}
+
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-g1} (using \texttt{scall} lemma)]
@@
           \]
-          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is defined as
-          \begin{align*}
-            \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
-            \phi_\pub & = \{(0,1)\}^* \\
-            \phi      & = (1,0) \union \phi_\pub \\
-            H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \uplus \{\npair[0]{\ms}\}
-          \end{align*}
-          % TODO show $\iota_x$ is a region.
-          show
+          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is the region in Definition~\ref{def:iotax-region} which is a region by Lemma~\ref{lem:iotax-is-a-region}.
+          Show
           \begin{equation}
@@
 \begin{lemma}
+  \label{lem:iotapwl-is-a-region}
+  $\iota^\pwl_{\start,\addrend}$ is a region for all $\start$ and $\addrend$.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{iotapwl-is-a-region}]
+  % TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:iotanwl-is-a-region}
+  $\iota^\nwl_{\start,\addrend}$ is a region for all $\start$ and $\addrend$.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{iotanwl-is-a-region}]
+  % TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:iotanwlp-is-a-region}
+  $\iota^{\nwl,p}_{\start,\addrend}$ is a region for all $\start$ and $\addrend$.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{iotanwlp-is-a-region}]
+  % TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:iotasta-is-a-region}
+  $\iota^\sta(v,\ms)$ is a region for all $v \in \{\perma, \temp\}$ and $\ms$.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{iotansta-is-a-region}]
+  % TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:iotastau-is-a-region}
+  $\iota^{\sta,u}(v,\ms)$ is a region for all $v \in \{\perma, \temp\}$ and $\ms$.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{iotanstau-is-a-region}]
+  % TODO
+\end{proof}
+
+\begin{lemma}
   \label{lem:revoketemp-readcond}

### 2017-02-02 local-capability-machine.tex

@@
 \end{mathpar}
+
+% XXX
+
+\begin{proof}[Proof of Theorem~\ref{thm:world-existence}]
+  The theorem follows from a more general solution theorem for the
+  category of $P$ preordered cofe's~\cite{bizjak-note}. 
+  We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to
+  $P$. 
+  \begin{align*}
+     F_1((X,{\futurestr}'),(Y,{\futurewk}')) = \\
+     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
+                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,{\futurewk}') \monnefun \UPred{\HeapSegments})) + \\
+                                          & \{\perma\} \times \States
+                                            \times \Rels \times
+                                            (\States \fun ((X,{\futurestr}')
+                                            \monnefun
+                                            \UPred{\HeapSegments})))),
+                                        \futurestr)
+  \end{align*}
+  and
+  \begin{align*}
+     F_2((X,{\futurestr}'),(Y,{\futurewk}')) = \\
+     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
+                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,{\futurewk}') \monnefun \UPred{\HeapSegments})) + \\
+                                          & \{\perma\} \times \States
+                                            \times \Rels \times
+                                            (\States \fun ((X,{\futurestr}')
+                                            \monnefun
+                                            \UPred{\HeapSegments})))),
+                                        \futurestr)
+  \end{align*}
+  The orderings $\futurestr$ and $\futurewk$ used in the definition of
+  $F_1$ and $F_2$ are defined by the properties given above.
+  Note that the image of $F_1$ and $F_2$ only differ in the ordering
+  relation, i.e., letting $U$ denote the forgetful functor from the
+  category of preordered cofe's to the category of cofe's, we have
+  $U \circ F_1 = U \circ F_2$.
+  From~\cite{bizjak-note} it then follows that there exists a cofe $\Wor$ and two
+  preorderings $\futurestr$ and $\futurewk$ and an isomorphism $\xi$ satisfying the properties
+  claimed in theorem.
+  (Here, in the proof, we have written the ordering explicitly on the cofe when
+  using monotone non-expansive functions; in the theorem formulation
+  we have instead annotated the arrow to indicate which ordering is used.)
+\end{proof}
+
+% XXX end
 

### 2017-02-02 local-capability-machine.tex
Finish the FTLR (modulo the lemmas factored out).

@@
   \item $\memheap = \biguplus_{r\in\activeReg{W}}P(r)$ and $\forall r \in
-    \activeReg{W}$, $n'<n \ldotp$ $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
-    and $\npair[n']{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
+    \activeReg{W}$, $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
+    and $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
     $\memSatPar{\memheap}{W}{P}$. \label{step:unfold-phi-sat-part}
@@
   \item $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$
-    and $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$.
+    and $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$.
     By specializing the result from
     Step~\ref{step:unfold-phi-sat-part} to the $r$ from
-    Step~\ref{step:unfold-readCond} and $n' = n-1$.
+    Step~\ref{step:unfold-readCond}.
+    \dominique{TODO: fix what follows for the recent step-related changes...}
   \item $\npair[n-1]{P(r)} \in H^\pwl_{\start',\addrend'}(s)(\xi^{-1}(W))$.
@@
    \begin{itemize}
-   \item $\heapSat[m]{n}{W}$
+   \item $\ms = \ms' \uplus \ms_f$
+   \item $\heapSat[\ms']{n}{W}$
    \item $((\perm,\gl),\start,\addrend,\addr) = c$
@@
 
-   then $\heapSat[{m\update{\addr}{\var{w}}}]{n}{W}$
+   then $\addr \in \dom(\ms')$ (i.e. $\ms\update{a}{w} =
+   \ms'\update{a}{w}\uplus\ms_f$) and
+   $\heapSat[{\ms'\update{\addr}{\var{w}}}]{n}{W}$
  \end{lemma}
@@
     By: definition of $\stder(W)$.
-  \item Assume: $\ms_f$, $\heap'$, $i \leq n'$, $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f) \step[i] (\halted,\Phi)$,\\
+  \item Assume: $\ms_f$, $\heap'$, $i \leq n'$, $\Phi =
+    (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)$
+    and $\Phi \step[i] (\halted,\Phi')$,\\
     Suffices: $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-    $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+    $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
     By: definition of $\observations(W)$ \label{suff-after-eval}
@@
     \end{equation*} \label{step:ip-non-zero}
-  \item $n' > 0$, since otherwise $i = 0$ (because $i \leq n'\leq n$) and this is
+  \item $n \geq n' > 0$, since otherwise $i = 0$ (because $i \leq n'\leq n$) and this is
     impossible by the previous point.
@@
   \item Case $\var{conf'} = \failed$: contradiction, since it is not possible
-    that $\failed \step[i'] (\halted,\Phi)$. 
+    that $\failed \step[i'] (\halted,\Phi')$. 
   \item Case $\var{conf'} = (\halted,\Phi)$:
@@
       \Phi$\\
-      Follows from $(\halted,
-      (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-      \ms_f)) \step[i'] (\halted,\Phi)$
+      Follows from $(\halted, \Phi) \step[i'] (\halted,\Phi)$
     \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
@@
     \begin{enumproof}
-    \item $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$.
+    \item $\memheap[\Phi''] = \ms'' \uplus \ms_f$ and $\heapSat[\ms'']{n'-1}{W}$.
       \begin{enumproof}
       \item Case $\memheap[\Phi''] = \memheap[\Phi]$:
-        $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
+        Then $\memheap[\Phi''] = \ms\uplus\ms_f$ and
+        $\heapSat[\ms]{n'-1}{W}$ follows by Lemma~\ref{lem:heap-sat-dc}.
       \item Case $\memheap[\Phi''] = \memheap[\Phi]\update{\addr'}{\var{w}}$,
@@
         \in \{\rwlx,\readwritel \}$.\\
-        $\heapSat[{\memheap[\Phi'']}]{n'-1}{W}$ follows by
-        Lemmas~\ref{lem:conds-store-suff} and~\ref{lem:heap-sat-dc} using the
-        fact that $\heapSat[\memheap]{n'}{W}$ and
+        The facts that $\memheap[\Phi''] = \ms''\uplus\ms_f$ and
+        $\heapSat[\ms'']{n'-1}{W}$ follow by Lemmas~\ref{lem:conds-store-suff}
+        and~\ref{lem:heap-sat-dc} using the fact that $\heapSat[\ms]{n'}{W}$ and
         $\npair[n']{\memreg(r_1)}\in\stdvr(W)$ and
         $\npair[n']{\memreg(r_2)}\in\stdvr(W)$ which follows from
-        $\npair[n']{\memreg}\in\stdrr(W)$ by definition.
+        Step~\ref{step:ftlr-reg-rr}.
       \end{enumproof}
@@
     \item
-      $\npair[n'-1]{(\memreg[\Phi'']\update{\pcreg}{\var{newPc}},\memheap[\Phi''])}\in\observations(W)$,
-      i.e. $\npair[n'-1]{\var{conf'}}\in\observations(W)$:
+      $\npair[n'-1]{(\memreg[\Phi'']\update{\pcreg}{\var{newPc}},\ms'')}\in\observations(W)$:
       by definition of $\stder(W)$ using the above three points.
     \item $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-      $\memheap[\Phi] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+      $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
       By: definition of $\observations(W)$ using the previous step and the
       evaluation $\var{conf'}\step[i'] (\halted,\Phi')$ from
-      Step~\ref{step:ip-non-zero}.\dominique{TODO: check.  What about $ms_f$ and other stuff.}
+      Step~\ref{step:ip-non-zero}.
     \end{enumproof}

### 2017-02-02 local-capability-machine.tex
Remove obsolete comment.
Remove the n-1 from readCond as discussed with Lau and use it to make the proof of Lemma lem:load-conds-suff go through.

@@
 \end{align*}
-\dominique{quantification over $n' < n$ needed?}
 
@@
     & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-    & \;\qquad W(r)\nsubsim[n-1] \iota_{\start',\addrend'}^\pwl \}
+    & \;\qquad W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl \}
   \end{aligned}
@@
   \item $\exists r \in \var{localityReg}(g,W)$, $[\start',\addrend'] \supseteq
-    [\start,\addrend] \ldotp$ $W(r)\nsubsim[n-1]
+    [\start,\addrend] \ldotp$ $W(r)\nsubsim[n]
     \iota_{\start',\addrend'}^\pwl$. By definition of $\readCond{\gl}(W)$. \label{step:unfold-readCond}
@@
     Step~\ref{step:unfold-readCond}.
-    \dominique{TODO: fix what follows for the recent step-related changes...}
-  \item $\npair[n-1]{P(r)} \in H^\pwl_{\start',\addrend'}(s)(\xi^{-1}(W))$.
-    Follows by combining $\npair[n-1]{P(r)} \in H(s)(\xi^{-1}(W))$ with
-    $W(r)\nsubsim[n-1] \iota_{\start',\addrend'}^\pwl$ from Step~\ref{step:unfold-readCond}.
+  \item $\npair[n]{P(r)} \in H^\pwl_{\start',\addrend'}(s)(\xi^{-1}(W))$.
+    Follows by combining $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$ with
+    $W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl$ from Step~\ref{step:unfold-readCond}.
   \item $\dom(P(r)) = [\start',\addrend']$ and for all $\addr' \in
     [\start',\addrend']\ldotp$
-    $\npair[n-2]{P(r)(\addr')}\in\stdvr(\xi(\xi^{-1}(W)))$.
+    $\npair[n-1]{P(r)(\addr')}\in\stdvr(\xi(\xi^{-1}(W)))$. By definition of $H^\pwl_{\start',\addrend'}$.
   \item $\addr \in [\start,\addrend]\subseteq[\start',\addrend']$. By combining
@@
   \item In particular, we get: $\memheap(\addr) = P(r)(\addr)$ and
-    $\npair[n-2]{P(r)(\addr)}\in\stdvr(W)$.
-    \dominique{TODO: Make sure we lose only one step in the definitions!}
+    $\npair[n-1]{P(r)(\addr)}\in\stdvr(W)$.
   \end{enumproof}

### 2017-02-02 local-capability-machine.tex
many changes to the region related proofs

@@
 \begin{proof}[Proof of Lemma~\ref{lem:iotax-is-a-region}]
-
+  \begin{itemize}
+  \item $\phi_\pub$ is defined as the reflexive transitive closure, so it is immediately well foremd.
+  \item $\phi$ adds a transition to $\phi_\pub$ and is also reflexive and transitive. 
+  \item $H_x$ is trivially non-expansive in the state.
+  \item $H_x$ does not depend on the $\hat{W}$, so it also becomes trivially non-expansive and (privately) monotone in $\hat{W}$.
+  \end{itemize}
 \end{proof}
@@
 \subsubsection{Standard regions}
+
+\begin{lemma}
+  \label{lem:hpwl-mono}
+  $H^\pwl_{\start,\addrend} \; s$ is monotone w.r.t $\futurewk$ for all $s \in \States$ and $\start$ and $\addrend$
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:hpwl-mono}]
+  Let $\hat{W}' \futurewk \hat{W}$ be given and let 
+  \begin{equation}
+    \label{pf:hpwl-mono:ass}
+    \npair{\ms} \in H^\pwl_{\start,\addrend}\; s \; \hat{W}
+  \end{equation}
+  and show
+  \[
+    \npair{\ms} \in H^\pwl_{\start,\addrend}\; s \; \hat{W}'
+  \]
+  From \ref{pf:hpwl-mono:ass}, we get $\dom(\ms) = [\start,\addrend]$. Now let $\addr \in [\start,\addrend]$ be given and show
+  \[
+    \npair[n-1]{\ms(a)} \in \stdvr(\xi(\hat{W}'))
+  \]
+  now this follows from Lemma~\ref{lem:stdvr-mono-wk}, $\hat{W}' \futurewk \hat{W}$, Theorem~\ref{thm:world-existence}, and Assumption~\ref{pf:hpwl-mono:ass}.
+\end{proof}
+
 \begin{lemma}
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{iotapwl-is-a-region}]
-  % TODO
+\begin{proof}[Proof of Lemma~\ref{lem:iotapwl-is-a-region}]
+  Follows from Lemma~\ref{lem:hpwl-mono}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:hnwl-mono-str}
+  $H^\nwl_{\start,\addrend} \; s$ is monotone w.r.t $\futurestr$ for all $s \in \States$ and $\start$ and $\addrend$
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:hnwl-mono-str}]
+  Let $\hat{W}' \futurewk \hat{W}$ be given and let 
+  \begin{equation}
+    \label{pf:hnwl-mono:ass}
+    \npair{\ms} \in H^\nwl_{\start,\addrend}\; s \; \hat{W}
+  \end{equation}
+  and show
+  \[
+    \npair{\ms} \in H^\nwl_{\start,\addrend}\; s \; \hat{W}'
+  \]
+  From \ref{pf:hnwl-mono:ass}, we get $\dom(\ms) = [\start,\addrend]$. Now let $\addr \in [\start,\addrend]$ be given and show
+  \[
+    \npair[n-1]{\ms(a)} \in \stdvr(\revokeTemp{\xi(\hat{W}')})
+  \]
+  %TODO
+  From Assumption~\ref{pf:hpwl-mono:ass}, we have
+  \[
+    \npair[n-1]{\ms(a)} \in \stdvr(\revokeTemp{\xi(\hat{W)}})
+  \]
+
+  now this follows from Lemma~\ref{lem:stdvr-mono-wk}, $\hat{W}' \futurewk \hat{W}$, Theorem~\ref{thm:world-existence}, and Assumption~\ref{pf:hpwl-mono:ass}.
 \end{proof}
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{iotanwl-is-a-region}]
+\begin{proof}[Proof of Lemma~\ref{lem:iotanwl-is-a-region}]
   % TODO
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{iotanwlp-is-a-region}]
+\begin{proof}[Proof of Lemma~\ref{lem:iotanwlp-is-a-region}]
   % TODO
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{iotansta-is-a-region}]
+\begin{proof}[Proof of Lemma~\ref{lem:iotasta-is-a-region}]
   % TODO
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{iotanstau-is-a-region}]
+\begin{proof}[Proof of Lemma~\ref{lem:iotastau-is-a-region}]
   % TODO
-\end{proof}
-
-\begin{lemma}
-  \label{lem:revoketemp-readcond}
-  If
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\gl)(\revokeTemp{W})
-  \]
-  then
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-readcond}]
-  \[
-    \npair{(\start,\addrend)} \in \readCond{}(\gl)(\revokeTemp{W})
-  \]
-  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
-  \[
-    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n-1] \iota^\pwl_{[\start',\addrend']}
-  \]
-  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:revoketemp-writecond}
-  If
-  \[
-    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
-  \]
-  then
-  \[
-    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
-  \[
-    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
-  \]
-  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
-  \[
-    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
-  \]
-  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:revoketemp-execcond}
-  If
-  \begin{itemize}
-  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(\revokeTemp{W})$
-  \end{itemize}
-  then
-  \[
-    \npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-execcond}]
-  Use Lemma~\ref{lem:rt-w-pub-future-w}.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:revoketemp-entercond}
-  If
-  \begin{itemize}
-  \item $\npair{(\addr,\start,\addrend)} \in \execCond{}(\gl)(\revokeTemp{W})$
-  \end{itemize}
-  then
-  \[
-    \npair{(\addr,\start,\addrend)} \in \execCond{}(\gl)(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-entercond}]
-    Use Lemma~\ref{lem:rt-w-pub-future-w}.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:revoketemp-vr-implies-vr}
-  If
-  \[
-    \npair{w} \in \stdvr(\revokeTemp{W})
-  \]
-  then
-  \[
-    \npair{w} \in \stdvr(W)
-  \]
-\end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-vr-implies-vr}]
-  Follows from Lemma~\ref{lem:revoketemp-readcond}, Lemma~\ref{lem:revoketemp-writecond}, Lemma~\ref{lem:revoketemp-execcond}, and Lemma~\ref{lem:revoketemp-entercond}.
 \end{proof}
@@
   \[
-    H^\nwl_{\start,\addrend} \; s\; \hat{W} \nsubseteq H^\pwl_{\start,\addrend} \; s\; \hat{W}
+    H^\nwl_{\start,\addrend} \; s\; \hat{W} \nsubeq H^\pwl_{\start,\addrend} \; s\; \hat{W}
   \]
@@
   \[
-    H^\nwl_{\start,\addrend} \; 1\; \hat{W} \nsubseteq H^\pwl_{\start,\addrend} \; 1\; \hat{W}  
+    H^\nwl_{\start,\addrend} \; 1\; \hat{W} \nsubeq H^\pwl_{\start,\addrend} \; 1\; \hat{W}  
   \]
@@
 \subsubsection{Permission based conditions}
+
+\begin{lemma}
+  \label{lem:revoketemp-readcond}
+  If
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(\revokeTemp{W})
+  \]
+  then
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-readcond}]
+  \[
+    \npair{(\start,\addrend)} \in \readCond{}(\gl)(\revokeTemp{W})
+  \]
+  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
+  \[
+    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n-1] \iota^\pwl_{[\start',\addrend']}
+  \]
+  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-writecond}
+  If
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
+  \]
+  then
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-writecond}]
+  \[
+    \npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(\revokeTemp{W})
+  \]
+  Gives $r \in \localityReg(\gl,\revokeTemp{W})$ such that
+  \[
+    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsupsim[n-1] \iota_{[\start',\addrend']}
+  \]
+  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-execcond}
+  If
+  \begin{itemize}
+  \item $\npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(\revokeTemp{W})$
+  \end{itemize}
+  then
+  \[
+    \npair{(\perm,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-execcond}]
+  Use Lemma~\ref{lem:rt-w-pub-future-w}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-entercond}
+  If
+  \begin{itemize}
+  \item $\npair{(\addr,\start,\addrend)} \in \execCond{}(\gl)(\revokeTemp{W})$
+  \end{itemize}
+  then
+  \[
+    \npair{(\addr,\start,\addrend)} \in \execCond{}(\gl)(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-entercond}]
+    Use Lemma~\ref{lem:rt-w-pub-future-w}.
+\end{proof}
+
 \begin{lemma}
@@
 \subsubsection{Value relation}
+\begin{lemma}[Value relation monotone wrt $\futurewk$]
+  \label{lem:stdvr-mono-wk}
+  \[
+    W' \futurewk W \land \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of lemma~\ref{lem:stdvr-mono-wk}]
+  Follows from Lemma~\ref{lem:readcond-mono-pub}, Lemma~\ref{lem:writecond-mono-pub}, Lemma~\ref{lem:execcond-mono-pub}, and Lemma~\ref{lem:entrycond-mono-pub}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:revoketemp-vr-implies-vr}
+  If
+  \[
+    \npair{w} \in \stdvr(\revokeTemp{W})
+  \]
+  then
+  \[
+    \npair{w} \in \stdvr(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoketemp-vr-implies-vr}]
+  Follows from Lemma~\ref{lem:revoketemp-readcond}, Lemma~\ref{lem:revoketemp-writecond}, Lemma~\ref{lem:revoketemp-execcond}, and Lemma~\ref{lem:revoketemp-entercond}.
+\end{proof}
+
 \begin{lemma}[Global capabilities monotone wrt $\futurestr$]
@@
   \end{enumproof}
+\end{proof}
+
+\begin{lemma}[Non local words monotone wrt $\futurestr$]
+  \label{lem:stdvr-non-loc-priv-mono}
+  \begin{align*}
+    & \forall n, \perm, \start, \addrend, \addr, W, W',w \ldotp \\
+    & \quad ( w = \stdcap[(\perm,\glob)] \lor w \in \ints) \land \\
+    & \quad \npair{w} \in \stdvr(W) \land W' \futurestr W \\
+    & \qquad \Rightarrow \npair{w} \in \stdvr(W') 
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stdvr-non-loc-priv-mono}]
+  If $w = \stdcap[(\perm,\glob)]$, then iet follows from Lemma~\ref{lem:stdvr-glob-priv-mono}.
+
+  If $w \in \ints$, then it follows from the fact that $i \in \stdvr(W'')$ for all $i \in \ints$ and $W'' \in \Worlds$.
 \end{proof}

### 2017-02-02 local-capability-machine.tex
Added non-local definition
iota^nwl updated to require non-local words
updated iota^stau to match updated definition of iota^nwl
changed some proofs to fit with the changes of last commit

@@
 \newcommand{\pointstostack}[3]{\ensuremath{#1} \text{ points to stack with } \ensuremath{#2} \text{ used and } \ensuremath{#3} \text{ unused}}
+\newcommand{\nonlocal}[1]{\ensuremath{#1} \text{ is non-local}}
 
@@
 They are otherwise the same as their non-''permit write local'' counterparts.
+
+Define the following predicate:
+\begin{definition}
+  \label{def:non-local-cap}
+  We say word $w$ "\nonlocal{w}" iff either
+  \begin{itemize}
+  \item $w = \stdcap{(\perm,\glob)}$ for some $\perm$, $\addr$, $\start$, and $\addrend$; or
+  \item $w \in \ints$
+  \end{itemize}
+\end{definition}
 
@@
                   \begin{itemize}
-                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}$.
+                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}$. %Lemma
                   \end{itemize}
@@
                       \begin{enumproof}
-                      \item Show $\forall a \in [\start_\adv,\addrend_\adv] \ldotp \npair[n'-1]{\ms(a)} \in \stdvr(\revokeTemp{W})$
-                      \item Notice: $W = \revokeTemp{W}$
-                      \end{enumproof}
+                      \item Show $\forall a \in [\start_\adv,\addrend_\adv] \ldotp \npair[n'-1]{\ms(a)} \in \stdvr(W)$
+                      \end{enumproof}
+                      (non-local requirement trivial - only capability if the link one, and it is $\glob$).
                       \begin{enumproof}
@@
           \]
-          Partitioning the memory segment in the components of the disjoint union, and given $n' < n$, the $\malloc$ part follows from assumption $\memSat{\ms_\malloc}{[0 \mapsto \iota_{\malloc,0}]}$ and the $\malloc$ specification.
+          Partitioning the memory segment in the components of the disjoint union, the $\malloc$ part follows from assumption $\memSat{\ms_\malloc}{[0 \mapsto \iota_{\malloc,0}]}$ and the $\malloc$ specification.
 
@@
           \[
-            \npair[n']{\ms_\link} \in H^{\sta,u}(\ms_\link)(1)(\xi^{-1}(W_{\var{mini}}))
+            \npair[n]{\ms_\link} \in H^{\sta,u}(\ms_\link)(1)(\xi^{-1}(W_{\var{mini}}))
           \]
@@
           \[
-            \npair[n'-1] \in \stdvr(W_{\var{mini}})
-          \]
-          which is indeed the case. \lau{We need to add a reason here! Is this part of the specification of malloc?}
+            \npair[n-1]{c_\malloc} \in \stdvr(W_{\var{mini}})
+          \]
+          which follows from Lemma~\ref{lem:malloc-in-vr}
 
@@
           \[
-            \npair[n']{\ms_\adv} \in H^\sta_{\start_\adv,\addrend_\adv}(1)(\xi^{-1}(W_{\var{mini}}))
+            \npair[n]{\ms_\adv} \in H^\sta_{\start_\adv,\addrend_\adv}(1)(\xi^{-1}(W_{\var{mini}}))
           \]
@@
           \[
-            \forall \addr \in \dom(\ms_\adv) \ldotp \npair[n'-1]{\ms_\adv(\addr)} \in \stdvr(W_{\var{mini}})
+            \forall \addr \in \dom(\ms_\adv) \ldotp \npair[n-1]{\ms_\adv(\addr)} \in \stdvr(W_{\var{mini}})
           \]
@@
           \[
-            \npair[n'-1]{c_\link} \in \stdvr(W_{\var{mini}})
+            \npair[n-1]{c_\link} \in \stdvr(W_{\var{mini}})
           \]
@@
           \[
-            \npair[n'-1]{(\link,\link)} \in \readCond{}(\glob)(W_{\var{mini}})
+            \npair[n-1]{(\link,\link)} \in \readCond{}(\glob)(W_{\var{mini}})
           \]
@@
           \[
-            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n'-2] \iota^\pwl_{\link,\link}
+            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-2] \iota^\pwl_{\link,\link}
           \]
@@
     \end{aligned}
-        \right\}
+        \right\} \union \{\npair[0]{\ms} \}
 \end{align*}
@@
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
+      &\quad \nonlocal{\ms(\addr)} \land\\ 
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
     \end{aligned}
-        \right\}
+        \right\} \union \{\npair[0]{\ms} \}
 \end{align*}
@@
   \iota^{\sta,u} (v,\ms) &= (v,1,=,=,H^{\sta,u}\;\ms)\\
-  H^\sta \; \ms \; s \; \hat{W} = & \left\{\npair{\ms'} \middle|
+  H^{\sta,u} \; \ms \; s \; \hat{W} = & \left\{\npair{\ms'} \middle|
     \begin{aligned}
-      &\ms' = \ms \\
-      &\forall \addr \in \dom(\ms) \ldotp \npair[n-1]{\ms(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
+      &\ms' = \ms \land\\
+      &\forall \addr \in \dom(\ms) \ldotp\\
+      & \quad \nonlocal{\ms(\addr)} \land\\
+      & \quad \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))
     \end{aligned}
-        \right\}
+        \right\} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
 \end{align*}
@@
   From \ref{pf:hnwl-mono:ass}, we get $\dom(\ms) = [\start,\addrend]$. Now let $\addr \in [\start,\addrend]$ be given and show
-  \[
-    \npair[n-1]{\ms(a)} \in \stdvr(\revokeTemp{\xi(\hat{W}')})
-  \]
-  %TODO
-  From Assumption~\ref{pf:hpwl-mono:ass}, we have
-  \[
-    \npair[n-1]{\ms(a)} \in \stdvr(\revokeTemp{\xi(\hat{W)}})
-  \]
-
-  now this follows from Lemma~\ref{lem:stdvr-mono-wk}, $\hat{W}' \futurewk \hat{W}$, Theorem~\ref{thm:world-existence}, and Assumption~\ref{pf:hpwl-mono:ass}.
+  \begin{enumproof}
+    \item \nonlocal{\ms(a)} \label{pf:hnwl-mono:ob1}
+    \item $\npair[n-1]{\ms(a)} \in \stdvr(\xi(\hat{W}'))$ \label{pf:hnwl-mono:ob2}
+  \end{enumproof}
+  \ref{pf:hnwl-mono:ob1} follows trivially from \label{pf:hnwl-mono:ass}.
+\ref{pf:hnwl-mono:ob1} follows from Assumption~\ref{pf:hpwl-mono:ass}, \ref{pf:hnwl-mono:ob1} (which we just argued), $W' \futurestr W$, and Lemma~\ref{lem:stdvr-non-loc-priv-mono}.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:iotanwl-is-a-region}]
-  % TODO
+  Follows from Lemma~\ref{lem:hnwl-mono-str} and Lemma~\ref{lem:future-pub-impl-future-priv}.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:iotanwlp-is-a-region}]
-  % TODO
+  Follows from Lemma~\ref{lem:hnwl-mono-str}.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:iotasta-is-a-region}]
-  % TODO
+  $H^\sta$ does not depend on $\hat{W}$, so it is trivial to show the necessary non-expansive and monotonicity requirements.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:hwnl-nsubset-hpwl}]
-  Let
+  Trivial. Let
   \[
@@
   \[
-    \npair[n-1]{\ms(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
-  \]
-  Now the desired reulst follows from Lemma~\ref{lem:revoketemp-vr-implies-vr}.
+    \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))
+  \]
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:nwlp-subset-pwl}]
-  Follows from Lemma~\ref{lem:hwnl-nsubset-hpwl}.
+  Follows from Lemma~\ref{lem:hwnl-nsubset-hpwl} (see proof of Lemma~\ref{lem:nwl-subset-pwl}).
 \end{proof}
@@
     & \forall n, \perm, \start, \addrend, \addr, W, W',w \ldotp \\
-    & \quad ( w = \stdcap[(\perm,\glob)] \lor w \in \ints) \land \\
+    & \quad \nonlocal{w} \land \\
     & \quad \npair{w} \in \stdvr(W) \land W' \futurestr W \\

### 2017-02-02 local-capability-machine.tex
added the last region proofs

@@
 \begin{proof}[Proof of Lemma~\ref{lem:hnwl-mono-str}]
-  Let $\hat{W}' \futurewk \hat{W}$ be given and let 
+  Let $\hat{W}' \futurestr \hat{W}$ be given and let 
   \begin{equation}
@@
   \end{enumproof}
-  \ref{pf:hnwl-mono:ob1} follows trivially from \label{pf:hnwl-mono:ass}.
-\ref{pf:hnwl-mono:ob1} follows from Assumption~\ref{pf:hpwl-mono:ass}, \ref{pf:hnwl-mono:ob1} (which we just argued), $W' \futurestr W$, and Lemma~\ref{lem:stdvr-non-loc-priv-mono}.
+  \ref{pf:hnwl-mono:ob1} follows trivially from \ref{pf:hnwl-mono:ass}.
+\ref{pf:hnwl-mono:ob1} follows from Assumption~\ref{pf:hpwl-mono:ass}, \ref{pf:hnwl-mono:ob1} (which we just argued), $\hat{W}' \futurestr \hat{W}$, Theorem~\ref{thm:world-existence}, and Lemma~\ref{lem:stdvr-non-loc-priv-mono}.
 \end{proof}
@@
 \begin{lemma}
+  \label{lem:hstau-mono-str}
+  $H^{\sta,u}(\ms) \; s$ is monotone w.r.t $\futurestr$ for all $s \in \States$ and $\ms$.
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:hstau-mono-str}]
+  Let $\hat{W}' \futurestr \hat{W}$ be given and let 
+  \begin{equation}
+    \label{pf:hstau-mono:ass}
+    \npair{\ms'} \in H^{\sta,u}(\ms)\; s \; \hat{W}
+  \end{equation}
+  and show
+  \[
+    \npair{\ms'} \in H^{\sta,u}(\ms)\; s \; \hat{W}'
+  \]
+  From \ref{pf:hstau-mono:ass}, we get $\ms' = \ms$. Now let $\addr \in \dom(\ms)$ be given and show
+  \begin{enumproof}
+    \item \nonlocal{\ms(a)} \label{pf:hstau-mono:ob1}
+    \item $\npair[n-1]{\ms(a)} \in \stdvr(\xi(\hat{W}'))$ \label{pf:hstau-mono:ob2}
+  \end{enumproof}
+  \ref{pf:hstau-mono:ob1} follows trivially from \ref{pf:hstau-mono:ass}.
+\ref{pf:hstau-mono:ob1} follows from Assumption~\ref{pf:hstau-mono:ass}, \ref{pf:hstau-mono:ob1} (which we just argued), $\hat{W}' \futurestr \hat{W}$, Theorem~\ref{thm:world-existence} and Lemma~\ref{lem:stdvr-non-loc-priv-mono}.
+\end{proof}
+
+\begin{lemma}
   \label{lem:iotastau-is-a-region}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:iotastau-is-a-region}]
-  % TODO
+  Follows from Lemma~\ref{lem:hstau-mono-str} and Lemma~\ref{lem:future-pub-impl-future-priv}.
 \end{proof}

### 2017-02-02 local-capability-machine.tex
Added lemmas about expression relation and so on

@@
 
+\subsubsection{Observation relation}
+\begin{lemma}[Observation relation ($\observations$) non-expansive]
+  \label{lem:obs-rel-ne}
+  \[
+    W \nequal W' \Rightarrow \observations(W) \nequal \observations(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:obs-rel-ne}]
+  % TODO
+\end{proof}
+
+\subsubsection{Register-file relation}
+\begin{lemma}[Register-file relation ($\stdrr$) non-expansive]
+  \label{lem:stdrr-ne}
+  \[
+    W \nequal W' \Rightarrow \stdrr(W) \nequal \stdrr(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stdrr-ne}]
+  % TODO
+\end{proof}
+
+\begin{lemma}[Register-file relation ($\stdrr$) monotone wrt $\futurewk$]
+  \label{lem:stdrr-mono}
+  \[
+    W' \futurewk W \Rightarrow \stdrr(W') \nsupeq \stdrr(W)
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stdrr-mono}]
+  % TODO
+\end{proof}
+
+\subsubsection{Expression relation}
+\begin{lemma}[Expression relation ($\stder$) non-exapansive]
+  \label{lem:stder-ne}
+  \[
+    W \nequal W' \Rightarrow \stder(W) \nequal \stder(W')
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:stder-ne}]
+
+\end{proof}
+
 \subsubsection{Permission based conditions}

### 2017-02-02 local-capability-machine.tex
Added comment about how logical relation is defined to be well-founded

@@
 \subsection{The logical relation}
-
+The logical relation is defined by several mutual recursive definitions.
+In order to handle this mutual recursion and show that this definitions are well-defined, Banach's fixed-point theorem can be used.
+% TODO: Reference?
+We have omitted the details of this construction here, but it is done by parameterising all the definitions by the value relation.
 \begin{align*}

### 2017-02-02 local-capability-machine.tex
spell check

@@
 \lau{At some point we talked about whether $\updatePcPerm{}$ should just give a $\glob$ capability all the time.
-I believe this is dangerous as it can be used to first get a $\glob$ read-capability for the stack and subsequently in a $\glob$ callback, you can start reading from someone elses private stack!}
+I believe this is dangerous as it can be used to first get a $\glob$ read-capability for the stack and subsequently in a $\glob$ callback, you can start reading from someone else's private stack!}
 % TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
@@
 
-\dominique{factor out reoccuring conditions of the form $n_i = \rv_i$ or $n_i =
+\dominique{factor out reoccurring conditions of the form $n_i = \rv_i$ or $n_i =
   \memreg(\rv_i)$ into a function that evaluates an $\rv$ using a system state?
-  This would generalize nicely to the case where we would have additional
+  This would generalise nicely to the case where we would have additional
   addressing modes.}
@@
 A $\mathtt{store}$ that allows integers to be stored directly.
-$\mathtt{store}$ requries a register $r_t$ for storage of temporary values to be available.
+$\mathtt{store}$ requires a register $r_t$ for storage of temporary values to be available.
 \begin{align*}
@@
 
-              The capability linking table can be shared between multiple functions that are linked to the same capabiblities as it is accessed through read-only capabilities.
+              The capability linking table can be shared between multiple functions that are linked to the same capabilities as it is accessed through read-only capabilities.
 
@@
 Restrict it to a local enter-capability and put it in $r_0$.
-\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?} \lau{It is also used when we want to call code for the first time, so calling it a protected return pointer is more of a conceptul thing.}
+\lau{Do we want to use ``protected return pointer'' to mean a local enter capability?} \lau{It is also used when we want to call code for the first time, so calling it a protected return pointer is more of a conceptual thing.}
                 \item [Restrict stack capability to unused part]
@@
                 \item[Store capabilities and activation code to record] Store the environment capability and code capability in the record followed by the activation code. 
-                \item[Restrict the capability for the ``clojure pair'' to an enter capability] Adjust the capability to point to the start of the activation code and restrict it to a global enter-capability.
+                \item[Restrict the capability for the ``closure pair'' to an enter capability] Adjust the capability to point to the start of the activation code and restrict it to a global enter-capability.
                 \item[Activation code:] \forcenewline
@@
               \item \texttt{reqperm} can be used to test whether something can pass as a stack.
-              \item \texttt{reqglob} can be used to test whether a callback is admissable in the presence of a stack.
+              \item \texttt{reqglob} can be used to test whether a callback is admissible in the presence of a stack.
               \item The code of a closure will often be found in conjunction with the code that creates it.
@@
                       The execution loads $c_l$ to $r_l$ and jumps to $c_{f1}^{(3)}$ (the point just before the assertion).
-As $\ms_l = 1$, the assetion is successful and the execution halts.
+As $\ms_l = 1$, the assertion is successful and the execution halts.
 In other words, there were no changes to the memory.\\
-                      Use $W'$, $\ms_r = \emptyset$, and $\ms'$ to get the desired result, i.e., $m' = \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n''-k''}{W'}$ (using downwardsclosure of memory satisfaction).
+                      Use $W'$, $\ms_r = \emptyset$, and $\ms'$ to get the desired result, i.e., $m' = \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n''-k''}{W'}$ (using downwards closure of memory satisfaction).
                     \end{enumproof}
@@
 \begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
-  Let $n$ be given and make the assunptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
+  Let $n$ be given and make the assumptions of the lemma. Consider the execution until just after the jump to $\adv$, i.e.,
   \begin{multline*}
@@
 The $\mathtt{assert}$ therefore succeeds making no changes to the memory and afterwards the execution halts.
-      \dominique{Maybe you can factor out some sort of reusable antireduction lemma
+      \dominique{Maybe you can factor out some sort of reusable anti-reduction lemma
         for this sort of argument?}
@@
 
-\lau{I have skipped the following program in favor of the one in the next section. }
+\lau{I have skipped the following program in favour of the one in the next section. }
               ML-like program:
@@
 At this point the top of the stack would be 2, so when the execution reaches the assertion, it would fail.
-However, the producted return pointer is passed as a local capability, so the only place the adversary can store it is on the stack.
+However, the produced return pointer is passed as a local capability, so the only place the adversary can store it is on the stack.
 The adversary loses control of the stack when control is returned to $f3$ where the $\mathtt{scall}$ makes sure to sanitise the stack and register file before control is passed back to the adversary.
@@
 
-To argue that the configuration is suitable, we need to argue that invoking the continuation produces an admissable result.
+To argue that the configuration is suitable, we need to argue that invoking the continuation produces an admissible result.
 As the continuation is a $\local$ capability, we take a public future world of $W$.
@@
 Due to Lemma~\ref{lem:scall-works} the \texttt{scall} lemma, for each \texttt{scall} we have to argue that the adversary and continuation produces results that respect the regions of $W$.
-Using Lemma~\ref{lem:anti-red-obs} the $\observations$ anti reduction lemma, it suffices to argue that each part of $f3$ between \texttt{scall}s produces admissable results.
+Using Lemma~\ref{lem:anti-red-obs} the $\observations$ anti reduction lemma, it suffices to argue that each part of $f3$ between \texttt{scall}s produces admissible results.
 
 Executing until the first \texttt{scall} only pushes 1 to the stack, so the invariants of $W$ are preserved.
-Due to the \texttt{scall} lemma, we need to argue that that the adversary and the continuation produce admissable results.
+Due to the \texttt{scall} lemma, we need to argue that that the adversary and the continuation produce admissible results.
 
@@
 As we provide no arguments to the adversary, most of the conditions are satisfied by assumptions and Lemma~\ref{lem:stack-cap-vr}, which makes sure that the stack capability is in the value relation.
-Which gives us that the adversary produces an admissable result.
+Which gives us that the adversary produces an admissible result.
 
@@
 Further, the program, flags, and linking table remain the same in any kind of future world.
-Therefore, we know that the execution continues by poping 1 from the stack and then asserting that it is indeed 1, which is indeed the case, so 2 is pushed to the stack.
+Therefore, we know that the execution continues by popping 1 from the stack and then asserting that it is indeed 1, which is indeed the case, so 2 is pushed to the stack.
 At this point we reach another scall.
@@
 The \texttt{scall} lemma lets us assume that the continuation continues in a memory that satisfies the invariants of $W$.
-The execution halts immediately in the continuation, so it produces an admissable result.
+The execution halts immediately in the continuation, so it produces an admissible result.
 \end{proof}
@@
       We know $W'(6)=\iota^\pwl_{\base_\stk'',\addrend_\stk}$ (using the definition of $W$ from the proof of Lemma~\ref{lem:correctness-f2}) which is a $\temp$ region.
-We futher know that $P(6) = \ms_{\var{?}}$ for some memory segment.
+We further know that $P(6) = \ms_{\var{?}}$ for some memory segment.
 By \ref{f3:mem-sat-par-mid} we know that $\npair[n'']{\ms_?} \in H^\pwl_{\base_\stk'',\addrend_\stk} \; 1 \; W'$ and thus $\ms_?
 = \ms_{\stk,\var{rest}}$.
-Sine $W'(6)$ is a $\temp$ region, we can from \ref{f3:ms-revoked} conlude that
+Sine $W'(6)$ is a $\temp$ region, we can from \ref{f3:ms-revoked} conclude that
       \begin{enumproof}[resume]
@@
 \subsection{Inverted control and return from closure}
-If we can prove the below example using the scall lemma, then we should also be able to prove the akward example.
+If we can prove the below example using the scall lemma, then we should also be able to prove the awkward example.
 \begin{lstlisting}
@@
   \begin{itemize}
-  \item $\phi_\pub$ is defined as the reflexive transitive closure, so it is immediately well foremd.
+  \item $\phi_\pub$ is defined as the reflexive transitive closure, so it is immediately well formed.
   \item $\phi$ adds a transition to $\phi_\pub$ and is also reflexive and transitive. 
@@
           If $\reg_1(r_0).\perm \in \{\entry, \exec, \rwx, \rwlx\}$, then either $\execCond{}$ or $\entryCond{}$ holds for the capability in $\reg_1(r_0)$.
-Now use $W_2 \futurewk W_1$ with the approriate condition to get
+Now use $W_2 \futurewk W_1$ with the appropriate condition to get
           \[
@@
                         \item $\npair[n_4-j-1]{\reg_5} \in \stdrr(W_5)$ where $\reg_5$ is as described in the \texttt{scall} lemma Hyp-callee premise.\\
-                          Amounts to showing:\lau{too deply nested for environments}\\
+                          Amounts to showing:\lau{too deeply nested for environments}\\
                           1) $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.
@@
                   \]
-                  and the fact that $\stdvr$ is monotone w.r.t $\futurewk$, $W_{10} \futurewk W_3$, and $\stdvr(W_{10})$ is downwardsclosed.
+                  and the fact that $\stdvr$ is monotone w.r.t $\futurewk$, $W_{10} \futurewk W_3$, and $\stdvr(W_{10})$ is downwards-closed.
                 \item otherwise \\
@@
             To this end the use $\reg_3(r_0) = \reg_9(r_0)$ and $\npair[n_4]{\reg_3(r_0)} \in \stdvr1(W_3)$.
-Assuming that $\reg_9(r_0).\perm \in \{\entry,\exec,\rwx,\rwlx \}$ (if this is not the case, then it is trivial to show the above as the execution fails), then either the $\execCond{}$ or the $\entryCond{}$ hold for approriate values.
+Assuming that $\reg_9(r_0).\perm \in \{\entry,\exec,\rwx,\rwlx \}$ (if this is not the case, then it is trivial to show the above as the execution fails), then either the $\execCond{}$ or the $\entryCond{}$ hold for appropriate values.
 Now use that $n_6-l < n_4$ and $W_{10} \futurewk W_3$ (\ref{g1:w10futwkw3})\footnote{We don't know whether the capability is local or global, but it does not matter as we have a public future world relation between the two worlds.} to get
@@
 We need a notion of subset between regions that is almost $n$-subset, but not quite.
-The only difference is that the view part of a region is diregarded.
+The only difference is that the view part of a region is disregarded.
 Define ``semi $n$-subset'' and ``semi $n$-supset'' as:
@@
 
-\subsubsection{Antireduction for the observation relation}
+\subsubsection{Anti-reduction for the observation relation}
 
@@
   \]
-  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immidiate.
+  Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immediate.
 \end{proof}
@@
   \]
-  to this end consider the possible cases of $\perm$ and show that each of the necesary conditions hold:
+  to this end consider the possible cases of $\perm$ and show that each of the necessary conditions hold:
   \begin{enumproof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:stdvr-non-loc-priv-mono}]
-  If $w = \stdcap[(\perm,\glob)]$, then iet follows from Lemma~\ref{lem:stdvr-glob-priv-mono}.
+  If $w = \stdcap[(\perm,\glob)]$, then let follows from Lemma~\ref{lem:stdvr-glob-priv-mono}.
 
@@
 programming language to be careful when handing out store-local capabilities:
-only the libc startup code should initialize the stack as store-local and malloc
+only the libc startup code should initialise the stack as store-local and malloc
 should \emph{not} produce them. This basically means that the libc
-initialization code (or whatever component produces the initial stack pointer)
+initialisation code (or whatever component produces the initial stack pointer)
 is part of our TCB.

### 2017-02-02 local-capability-machine.tex
comment with old attempt at g1 proof removed

@@
     \phi      & = (1,0) \union \phi_\pub \\
-    H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \uplus \{\npair[0]{\ms}\}
+    H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \union \{\npair[0]{\ms}\}
   \end{align*}
@@
 \end{lemma}
-\begin{proof}[Proof of Lemma~\ref{lem:iotax-is-a-region}]
+\begin{proof}[Proof of Lemma~\ref{lem:iotax-is-a-region}] \forcenewline
   \begin{itemize}
@@
           \]
-          %TODO elaborate on what happens here using anti reduction lemma and malloc spec.
+          % TODO elaborate on what happens here using anti reduction lemma and malloc spec.
           Use Lemma~\ref{lem:anti-red-obs}, so let $\ms_f'$ be given, then
@@
 \end{proof}
-
-\begin{comment}
-\begin{proof}[Proof of Lemma~\ref{lem:correctness-g1}]
-  Let $n$ be given and make the remaining assumptions of the lemma. Define
-  \begin{align*}
-    W = & [0 \mapsto \iota^\sta (\perma,\ms_{g1})]\\
-        & [1 \mapsto \iota^\sta (\perma,\ms_\flag)]\\
-        & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
-        & [3 \mapsto \iota^\pwl_{\start_\stk, \addrend_\stk}]\\ 
-        & [4 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}] \\
-        & [5 \mapsto \iota_{\malloc,0}]
-  \end{align*}
-  \[
-    \ms = \ms_{g1} \uplus 
-          \ms_\flag \uplus                
-          \ms_\link \uplus 
-          \ms_\adv \uplus 
-          \ms_\malloc \uplus 
-          \ms_\stk
-  \]
-  and use the FTLR on $c_\adv$ (possible due to lemma~\ref{lem:nwlp-subset-pwl} and the chosen world). From this, we get
-  \[
-    \npair[n+1]{c_\adv} \in \stder(W).
-  \]
-  Assuming we can show $\npair[n+1]{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$ and $\memSat[n+1]{\ms}{W}$, then we know
-  \[
-    \npair[n+1]{(\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}\update{\pcreg}{c_\adv},\ms) \in \observations(W)}
-  \]
-  Using $\ms_{\var{frame}}$ and resulting memory $m$ and the assumption from the lemma that
-  \[
-    (\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}\update{\pcreg}{c_\adv},m) \step[n] (\halted,m')
-  \]
-  We get some $W_{\var{final}} \futurestr W$, $\ms'$, and $\ms_r$ such that
-  \begin{align*}
-    & m' = \ms' \uplus \ms_r \uplus \ms_{\var{frame}}\\
-    & \memSat[1]{\ms'}{W_{\var{final}}}
-  \end{align*}
-  From island 1 (the flag island), the above memory segment satisfaction, and the fact that $W_{\var{final}} \futurestr W$ we get that part of $\ms'$ must be $\ms_\flag$ and thus
-  \[
-    m'(\flag) = 0
-  \]
-  which is the desired result. Now it suffices to show:
-  \begin{enumproof}
-  \item $\memSat[n+1]{\ms}{W}$
-    % TODO - this should be similar to previous proofs
-  \item $\npair[n+1]{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$
-    \begin{enumproof}
-    \item Case $r \in \RegName \setminus \{\pcreg,r_\stk,r_1,\}$: amounts to $\npair[n+1]{0} \in \stdvr(W)$, trivial.
-    \item Case $\npair[n+1]{c_\stk} \in \stdvr(W)$:
-      % TODO - same as in previous proof.
-    \item Case $\npair[n+1]{c_{g1}} \in \stdvr(W)$:\\
-      let $n' < n+1$ and $W' \futurestr W$ be given and show
-      \[
-        \npair[n']{\updatePcPerm{c_{g1}}} \in \stder(W')
-      \]
-      to this end let $n'' \leq n'$, $\ms'$, and $\reg'$ be given and assume
-      \begin{align}
-        & \npair[n'']{\reg'} \in \stdrr(W') \label{g1:regrel1} \\
-        & \memSat[n'']{\ms'}{W'} \label{g1:memsat1}
-      \end{align}
-      and show
-      \[
-        \npair[n'']{(\reg'\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms')} \in \observations(W')
-      \]
-      To this end let $i \leq n''$, $\ms_f$, $m''$ be given and assume
-      \begin{equation}
-        (\reg'\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms' \uplus \ms_f) \step[i] (\halted,m'') \label{g1:execution}
-      \end{equation}
-      and show
-      \begin{equation}
-        \exists W_{\var{str}}' \futurestr W', ms_r, \ms'' \ldotp m'' = \ms'' \uplus \ms_r \uplus \ms_f \land \memSat[n''-i]{\ms''}{W_{\var{str}}'} \label{g1:g-goal}
-      \end{equation}
-      consider how the above execution goes. By \ref{g1:memsat1} we know that the instructions of $g1$ are unaltered. The following happens
-      \begin{itemize}
-      \item ...
-        % TODO - typeout the details. 
-      \end{itemize}
-      In the end, we have a new memory segment
-      \[
-        \ms'' = \ms_\malloc' \uplus \ms_x \uplus \ms_{\var{env}} \uplus \ms_{\var{cls}}
-      \]
-      where $\ms_x(x) = 0$, $\ms_{\var{env}}(\var{env}) = ((\rwx,\glob),x,x,x)$, and $\ms_{\var{cls}}=[\var{cls} \mapsto c_{\var{env}}, \var{cls} + 1 \mapsto c_{f4}, \var{cls} + 2 \mapsto i^{ac}_1, \var{cls} + m + 1 \mapsto i^{ac}_m]$ where $i^{ac}_1, \dots, i^{ac}_m$ are the instructions of the activation code, $c_{\var{env}} = ((\rwx,\glob),\var{env},\var{env},\var{env})$, $c_{f4} = ((\exec,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{f4})$.
-      Further at this point, the register file contains:
-      \[
-        \reg'' = reg_0\update{\pcreg}{\updatePcPerm{\reg'(r_0)}}\update{r_0}{\reg'(r_0)}\update{r_1}{((\entry,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)}
-      \]
-      We know that $\npair[n'']{\reg'(r_0)} \in \stdvr(W')$.
-We have assumed execution does not fail, so $\reg'(r_0)$ must be a capability with permission $\entry$, $\exec$, $\rwx$, or $\rwlx$.
-In either case, we get $\execCond{}$ or $\entryCond{}$ (for certain parameters).
-Now use $n^{(3)}$ ($n''$ minus the steps the two malloc calls used) and $W'' \futurewk W'$ where
-      \[
-        W'' = W'\update{5}{\iota_\malloc'}
-        \update{i_1}{\iota^\sta (\perm,\ms_{\var{env}})}
-        \update{i_2}{\iota^\sta (\perm,\ms_{\var{cls}})}
-        \update{i_3}{\iota_{x,0}}
-      \]
-      where $i_1,i_2,i_3 \not\in\dom(W')$ and they distinct, and $\iota_\malloc \futurewk W(5)$ and
-      \begin{align*}
-        \iota_{x,s} & = (\perm,s,\{(0,1),(0,0),(1,1)\},\{(1,0),(0,1),(0,0),(1,1)\},H_x) \\
-        H_x \; s \; W & = \{\npair{\ms} \mid \ms(x) = s \land \dom(\ms) = [x,x]\}
-      \end{align*}
-      % TODO insert tikz picture of machine (just like Dreyer Neis machine.) and prove that it is indeed a region (should be simple, interpretation does not depend on world).
-      with $\execCond{}$ or $\entryCond{}$ to get
-      \[
-        \npair[n^{(3)}]{\updatePcPerm{\reg'(r_0)}} \in \stder(W'')
-      \]
-      If we can show $\npair[n^{(3)}]{\reg''} \in \stdrr(W'')$ and $\memSat[n^{(3)}]{\ms''}$, then using assumption \ref{g1:execution} we can get \ref{g1:g-goal}.
-
-      SFTS:
-      \begin{enumproof}
-      \item$\memSat[n^{(3)}]{\ms''}$
-        % TODO should be simple - all existing regions have interpretations monotone wrt public future worlds.
-The malloc island uses its "isolation condition" which makes room for the two new islands.
-The x islands starts in state 0 (and the appropriate memory address contains 0).
-The env and cls islands are static islands designed to accept exactly what the memory contains.
-      \item $\npair[n^{(3)}]{\reg''} \in \stdrr(W'')$ amounts to showing:
-        \begin{enumproof}
-        \item For $r \in \RegName \setminus \{\pcreg, r_0, r_1 \}$ $\reg''(r) = 0$ and it is trivial to show $\npair[n^{(3)}]{0} \in \stdvr(W'')$. 
-        \item $\npair[n^{(3)}]{\reg'(r_0)} \in \stdvr(W'')$ - follows from \ref{g1:regrel1} and public monotonicity of $\stdvr$.
-        \item $\npair[n^{(3)}]{((\entry,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)} \in \stdvr(W'')$ \\
-          To this end let $n^{(4)} < n^{(3)}$, $W^{(3)} \futurestr W''$ be given and assume
-          \begin{align}
-            \\ \npair[n^{(4)}]{\reg^{(3)}} \in \stdrr(W^{(3)}) \label{g1:reg-file-cls} &
-            \\ \memSat[n^{(4)}]{\ms^{(3)}}{W^{(3)}} \label{g1:memsat-cls}
-          \end{align}
-          for some $\reg^{(3)}$ and $\ms^{(3)}$ and show
-          \[
-            \npair[n^{(4)}]{(\reg^{(3)}\update{\pcreg}{((\exec,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)},\ms^{(3)})} \in \observations(W^{(3)})
-          \]
-          to this end let $j \leq n^{(4)}$, $m^{(3)}$, and $m_f'$ be given and assume
-          \begin{equation}
-            (\reg^{(3)}\update{\pcreg}{((\exec,\glob),\var{cls},\var{cls} + m + 1,\var{cls}+2)},\ms^{(3)} \uplus \ms_f') \step[j] (\halted, m^{(3)}) \label{g1:execution2}
-          \end{equation}
-          and show
-          \begin{equation}
-            \exists W_{\var{str}}^{(3)} \futurestr W^{(3)}, \ms'^{(3)}, \ms_r \ldotp m^{(3)} = \ms'^{(3)} \uplus \ms_r \uplus \ms_f' \land \memSat[n^{(4)}-j]{\ms'^{(3)}}{W_{\var{str}}'} \label{g1:g-goal2}
-          \end{equation}
-          to this end consider what happens in execution \ref{g1:execution2}: Due to \label{g1:memsat-cls} we know that the activation code and $f4$ code is unaltered.
-Running the activation code loads the environment capability to $r_{\var{env}}$ and loads the $f4$ capability, then the $r_1$ register is tested whether it contains a global capability (which it does s execution does not fail) and the $r_\stk$ register is tested for whether it contains an $\rwlx$ capability (which it does as execution does not fail), and the pointer address is adjusted, so the $r_\stk$-register now contains $((\rwlx, \gl), \start_{\stk}', \addrend_{\stk}', \start_{\stk}'-1)$.
-
-          Continuing execution until just after the point where we jump to the capability in register $r_1$ will store $0$ to address $x$ of memory, push restore code, $\reg^{(3)}(r_0)$, $\reg^{(3)}(r_1)$, push $\reg^{(3)}(r_\stk)$, and a capability to return to in $f4$.
-
-          Now use \ref{g1:memsat-cls} to get $P$ such that $\memSat[n^{(4)}]{\ms^{(3)}}{W^{(3)}}{P}$. Now use Lemma~\ref{lem:priv-mono-like2} to get $\ms^{(4)}$ and $\ms_R$ such that
-          \begin{align*}
-            & \ms^{(3)} = \ms^{(4)} \uplus \ms_R \\
-            & \memSat[n^{(4)}]{\ms^{(4)}}{\revokeTemp{W^{(3)}}}{P|_{\dom(\erase{W^{(3)}}{\perma})}} \\
-            & \ms_R = \biguplus_{r \in \erase{W^{(3)}}{\temp}} P(r) \land \\
-            & \ms^{(4)} = \biguplus_{r \in \erase{W^{(3)}}{\perma}} P(r)
-          \end{align*}
-          Now we want to argue that the ``stack'' we received is part of the revoked part.
-To do so, it suffices to show that it is governed by a temporary region.
-From \ref{g1:reg-file-cls} and $\reg^{(3)}(r_\stk)$ is a $\rwlx$ capability, we get
-          \begin{align*}
-            &\npair[n^{(4)}]{\start_\stk',\addrend_\stk'} \in \readCond{}(\gl)(W^{(3)}) \\
-            &\npair[n^{(4)}]{\start_\stk',\addrend_\stk'} \in \writeCond{}(\iota^\pwl,\gl)(W^{(3)})
-          \end{align*}
-          from which we can infer that for some $\start_\stk''$, $\addrend_\stk'$, and $r\in\dom(W^{(3)})$
-          \[
-            W^{(3)}(r) \nsim[n^{(4)}-1] \iota^\pwl_{\start_\stk'',\addrend_\stk''}
-          \]
-          from the fact that execution does not fail, we can infer $\start_\stk'' \leq \addrend_\stk''$. Now use Lemma~\ref{lem:pwl-nsim-view} to get
-          \[
-            W^{(3)}(r) \nequal[n^{(4)}-1] \iota^\pwl_{\start_\stk'',\addrend_\stk''}
-          \]
-          from which we can infer that the view of $W^{(3)}(r)$ is $\temp$.
-
-          We can from this infer that $\ms^{(3)} |_{\start_\stk'',\addrend_\stk''} \subset \ms_R$.
-
-          Consider the changes made to memory: $x$ was set to 0, so we have $\ms^{(4)}[x \mapsto 0]$.$W^{(3)}(i_3) \futurestr W''(i_3)$, so $W^{(3)}(i_3) \in \{0,1\}$.
-Now define $W^{(3)}_0 = \revokeTemp{W^{(3)}}[i_3 \mapsto \iota_{x,0}]$ we have $W^{(3)}_0 \futurestr W^{(3)}$.
-Further
-          \[
-            \memSat[n^{(4)}]{\ms^{(4)}[x \mapsto 0]}{W^{(3)}_0}{P|_{\dom(\erase{W^{(3)}}{\perma})}[i_3 \mapsto [x \mapsto 0]]}
-          \]
-          which follows from $H$ of $\perma$ regions being mono wrt $\futurestr$ and $x$ mapping to 0.
-
-          Define 
-          \[
-            W_{\var{reused}} = [i_1' \mapsto \iota^\sta (\temp,\ms|_{\start_\stk'',\start_\stk'''-1})][i_2' \mapsto \iota^\pwl_{\start_\stk''',\addrend_\stk''}]
-          \]
-          for $i_1',i_2' \not\in \dom(W^{(3)})$ and distinct where $\start_\stk'''$ is the first address of the stack we are passing on.
-
-          It is the case that
-
-          % Stack and x are the only things to change in memory.
-
-          % TODO: continue here.
-
-          The memory and register file at this point are as follows: %TODO
-          \[
-            \dots  
-          \]
-
-          We have assumed that the execution does not fail which entails several things: The capability in $r_1$ is executable after being jumped to.
-The capability $r_\stk$ has authority over some none empty part of the memory.
-
-          The latter is used with Lemma~\ref{lem:pwl-nsim-view}
-
-          As $\reg^{(3)}(r_\stk) = ((\rwlx,\gl),\start,\addrend,\addr)$ for some $\gl$, $\start$, $\addrend$, and $\addr$ and $\npair[n^{(4)}]{\reg^{(3)}(r_\stk)} \in \stdvr(W^{(3)})$ we know that
-          \begin{align*}
-            \\ \npair[n^{(4)}]{(\start,\addrend)} \in \readCond{}(\gl)(W) &
-            \\ \npair[n^{(4)}]{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
-          \end{align*}
-          which means that for some region $r$ and interval $[\start',\addrend'] \supseteq [\start,\addrend]$ we have 
-          \[
-            W(r) \nsim[n^{(4)}] \iota^\pwl_{\start',\addrend'}
-          \]
-
-        \end{enumproof}
-      \end{enumproof}
-    \end{enumproof}
-  \end{enumproof}
-\end{proof}
-
-\lau{If $c_{g1}$ is global it seems like we will run into trouble. When we have to prove $\npair{c_{g1}} \in \stdvr(W)$ - for now g1 is global (then we will see if it gives any troubles later on)}
-\end{comment}
 

### 2017-02-02 local-capability-machine.tex
Update a bunch of nsubsims to adapt to the change in readCond.

--- 2017-02-02 local-capability-machine.tex
@@
                   \begin{itemize}
-                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}$. %Lemma
+                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}
+                    \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: see Lemma~\ref{lem:nwlp-subset-pwl}.
                   \end{itemize}
@@
                       \item $a = \start_\adv$: show $((\readonly,\glob),\link,\link+1,\link) \in \stdvr(W)$\\
-                        SFTS $\iota_{\link} \nsubsim[n'-1] \iota^\pwl_{\link,\link+1}$\\
+                        SFTS $\iota_{\link} \nsubsim[n'] \iota^\pwl_{\link,\link+1}$\\
                         $\iota_{\link}$ accepts only $\ms_\link$, so show:
                         \begin{enumproof}
-                        \item $\npair[n'-2]{c_\malloc} \in \stdvr(W)$: We assume this is indeed the case.
-                        \item $\npair[n'-2]{c_\adv} \in \stdvr(W)$: for $n'' < n'-2$ and $W' \futurestr W$ show: \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stdvr(W')$. Use FTLR. \label{f1:adv}
+                        \item $\npair[n'-1]{c_\malloc} \in \stdvr(W)$: We assume
+                          this is indeed the case.\dominique{TODO: prove}
+                        \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: for $n'' <
+                          n'-1$ and $W' \futurestr W$ show:
+                          \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stdvr(W')$.
+                          Use FTLR.\dominique{no: by induction hypothesis?} \label{f1:adv}
                         \end{enumproof}
@@
   \[
-    \iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}
+    \iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}
   \]
@@
         \[
-          \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-        \]
-        which is indeed the case.
+          \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n] \iota^\pwl_{\start_\stk'', \addrend_\stk}
+        \]
+        which is trivially the case.
       \item $\npair{(\start_\stk'', \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$\\
@@
           \[
-            \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-          \]
-          which is indeed the case.
+            \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'] \iota^\pwl_{\start_\stk'', \addrend_\stk}
+          \]
+          which is trivially the case.
         \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \writeCond{}(\local)(W')$ \\
@@
           \[
-            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-2] \iota^\pwl_{\link,\link}
+            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-1] \iota^\pwl_{\link,\link}
           \]
@@
   \end{align*}
+  \dominique{err.. shouldn't you require that $\npair{\ms(a)}\in\stdvr(W)$ for
+    all $W$.  Hm: this probably doesn't hold like that?}
 \end{lemma}
@@
   \[
-    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n-1] \iota^\pwl_{[\start',\addrend']}
+    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n] \iota^\pwl_{[\start',\addrend']}
   \]

### 2017-02-02 local-capability-machine.tex
Prove the lemma about restrict.

@@
 
+ \begin{lemma}[pwl writecond implies nwl]
+   \label{lem:pwl-writecond-implies-nwl}
+   If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
+   $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$.
+ \end{lemma}
+ \begin{proof}
+   \begin{enumproof}
+   \item $\exists r \in \var{localityReg}(g,W) \ldotp$ $\exists
+     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
+     \iota^\pwl_{\start',\addrend'}$: by definition of $\writeCond{}$.
+   \item Suffices: $W(r)\nsupsim[n-1] \iota^\nwl_{\start',\addrend'}$. By
+     definition of $\writeCond{}$
+   \item $W(r)\nsupsim[n-1] \iota^\pwl_{\start',\addrend'} \nsupsim[n-1]
+     \iota^\nwl_{\start',\addrend'}$: follows by Lemma~\ref{lem:nwl-subset-pwl}.
+   \end{enumproof}
+ \end{proof}
+
  \begin{lemma}[Conditions for restrict instruction are sufficient]
@@
  \begin{proof}
-   TODO
+   By inspection of the definition of $\stdvr(W)$, everything follows trivially
+   except the result from lemma~\ref{lem:pwl-writecond-implies-nwl}.
+
+    If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$
+
+  \end{enumproof}
  \end{proof}

### 2017-02-02 local-capability-machine.tex
Fix a latex problem I caused.
Oops, forgot a case.

@@
 
+ \begin{lemma}[execCond implies entryCond]
+   \label{lem:execCond-implies-entryCond}
+   If $\npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W)$ then
+     $\npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)$.
+ \end{lemma}
+ \begin{proof}
+   \begin{enumproof}
+   \item Assume: $n' < n$, $W' \future W$ where $\gl = \local \Rightarrow
+     \future = \futurewk$ and $\gl = \glob \Rightarrow \future = \futurestr$
+     \\
+     Suffices: $\npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')$
+   \item Case $\addr \in [\start,\addrend]$: Follows from the definition of
+     $\execCond{}$.
+   \item Case $\addr \not\in[\start,\addrend]$: Follows by
+     Lemma~\ref{lem:failed-obs-stder}.
+   \end{enumproof}
+ \end{proof}
+
  \begin{lemma}[Conditions for restrict instruction are sufficient]
@@
    By inspection of the definition of $\stdvr(W)$, everything follows trivially
-   except the result from lemma~\ref{lem:pwl-writecond-implies-nwl}.
-
-    If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$
-
+   except the following.
+
+   \begin{enumproof}
+   \item If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
+     $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)$: holds
+     by lemma~\ref{lem:pwl-writecond-implies-nwl}.
+   \item If $\npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W)$ then
+     $\npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)$.
    \end{enumproof}

### 2017-02-02 local-capability-machine.tex
Prove another of the FTLR lemmas and fix a bug in the semantics: subseg should not be allowed for enter caps.
bib, proof

@@
                                               \begin{array}[t]{l}
-                                                \text{if $\memreg(r) = \stdcap[\permp]$} \\
+                                                \text{if $\memreg(r) = \stdcap$} \\
                                                 \quad\text{and for $i \in \{1,2\}$}\\
@@
                                                 \quad\text{and $\start \leq n_1$ and $n_2 \leq \addrend$}\\
-                                                \quad\text{and $c = (\permp,n_1,n_2,\addr)$}
+                                                \quad\text{and $\perm \neq \entry$}\\
+                                                \quad\text{and $c = ((\perm,\gl),n_1,n_2,\addr)$}
                                               \end{array} \\
@@
   The theorem follows from a more general solution theorem for the
-  category of $P$ preordered cofe's~\cite{bizjak-note}. 
+  category of $P$ preordered \cofes, see~\cite{bizjak-note}. 
   We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to
@@
   relation, i.e., letting $U$ denote the forgetful functor from the
-  category of preordered cofe's to the category of cofe's, we have
+  category of preordered \cofes\ to the category of \cofes, we have
   $U \circ F_1 = U \circ F_2$.
-  From~\cite{bizjak-note} it then follows that there exists a cofe $\Wor$ and two
+  From~\cite{bizjak-note} it then follows that there exists a \cofe\ $\Wor$ and two
   preorderings $\futurestr$ and $\futurewk$ and an isomorphism $\xi$ satisfying the properties
   claimed in theorem.
-  (Here, in the proof, we have written the ordering explicitly on the cofe when
+  (Here, in the proof, we have written the ordering explicitly on the \cofe\ when
   using monotone non-expansive functions; in the theorem formulation
@@
    \begin{itemize}
-   \item $\npair{((\perm',\gl'),\start',\addrend',\addr')} \in \stdvr(W)$
-   \item $\start' \leq \start''$
-   \item $\addrend'' \leq \addrend'$
-   \end{itemize}
-
-   then $\npair{((\perm',\gl'),\start'',\addrend'',\addr')} \in \stdvr(W)$
+   \item $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stdvr(W)$
+   \item $\start \leq \start'$
+   \item $\addrend' \leq \addrend$
+   \item $\perm \neq \entry$
+   \end{itemize}
+
+   then $\npair{((\perm,\gl),\start',\addrend',\addr)} \in \stdvr(W)$
  \end{lemma}
  \begin{proof}
-   TODO
+   Follows easily from the definitions of $\stdvr(W)$, $\readCond{}$,
+   $\writeCond{}$, $\execCond{}$.
  \end{proof}
@@
       $\addrend'' \leq \addrend'$ and $c =
-      ((\perm',\gl'),\start'',\addrend'',\addr')$
+      ((\perm',\gl'),\start'',\addrend'',\addr')$ and $\perm' \neq \entry$
     \end{itemize}
@@
         $\addrend'' \leq \addrend'$ and $c =
-        ((\perm',\gl'),\start'',\addrend'',\addr')$: \\
+        ((\perm',\gl'),\start'',\addrend'',\addr')$ and $\perm' \neq \entry$: \\
         $\npair[n'-1]{\memreg[\Phi''](r)} \in \stdvr(W)$ follows by

### 2017-02-03 local-capability-machine.tex
A few fixes to awkward example proof (overlooked in previous change)

--- 2017-02-02 local-capability-machine.tex
@@
                   \begin{itemize}
-                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}
-                    \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: see Lemma~\ref{lem:nwlp-subset-pwl}.
+                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}$. %Lemma
                   \end{itemize}
@@
                       \item $a = \start_\adv$: show $((\readonly,\glob),\link,\link+1,\link) \in \stdvr(W)$\\
-                        SFTS $\iota_{\link} \nsubsim[n'] \iota^\pwl_{\link,\link+1}$\\
+                        SFTS $\iota_{\link} \nsubsim[n'-1] \iota^\pwl_{\link,\link+1}$\\
                         $\iota_{\link}$ accepts only $\ms_\link$, so show:
                         \begin{enumproof}
-                        \item $\npair[n'-1]{c_\malloc} \in \stdvr(W)$: We assume
-                          this is indeed the case.\dominique{TODO: prove}
-                        \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: for $n'' <
-                          n'-1$ and $W' \futurestr W$ show:
-                          \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stdvr(W')$.
-                          Use FTLR.\dominique{no: by induction hypothesis?} \label{f1:adv}
+                        \item $\npair[n'-2]{c_\malloc} \in \stdvr(W)$: We assume this is indeed the case.
+                        \item $\npair[n'-2]{c_\adv} \in \stdvr(W)$: for $n'' < n'-2$ and $W' \futurestr W$ show: \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stdvr(W')$. Use FTLR. \label{f1:adv}
                         \end{enumproof}
@@
   \[
-    \iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}
+    \iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n-1] \iota^\pwl_{\start_\adv,\addrend_\adv}
   \]
@@
         \[
-          \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-        \]
-        which is trivially the case.
+          \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
+        \]
+        which is indeed the case.
       \item $\npair{(\start_\stk'', \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$\\
@@
           \[
-            \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'] \iota^\pwl_{\start_\stk'', \addrend_\stk}
-          \]
-          which is trivially the case.
+            \iota^\pwl_{\start_\stk'', \addrend_\stk} \nsubsim[n'-1] \iota^\pwl_{\start_\stk'', \addrend_\stk}
+          \]
+          which is indeed the case.
         \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \writeCond{}(\local)(W')$ \\
@@
           \[
-            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-1] \iota^\pwl_{\link,\link}
+            \iota^{\sta,u} (\perma,\ms_\link) \nsubsim[n-2] \iota^\pwl_{\link,\link}
           \]
@@
           \begin{itemize}
-          \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{4}{\iota_\malloc}} $
+          \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{0}{\iota_\malloc}} $
           \item $\iota_\malloc \futurewk \iota_{\malloc,0}$
@@
           \[
-            W_2 = W_1\update{4}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\update{i_2}{\iota_x}
+            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\update{i_2}{\iota_x}
           \]
@@
   \end{align*}
-  \dominique{err.. shouldn't you require that $\npair{\ms(a)}\in\stdvr(W)$ for
-    all $W$.  Hm: this probably doesn't hold like that?}
 \end{lemma}
@@
   \[
-    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n] \iota^\pwl_{[\start',\addrend']}
+    \forall [\start',\addrend'] \subseteq [\start,\addrend] \ldotp \revokeTemp{W}(r) \nsubsim[n-1] \iota^\pwl_{[\start',\addrend']}
   \]
@@
 
- \begin{lemma}[pwl writecond implies nwl]
-   \label{lem:pwl-writecond-implies-nwl}
-   If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
-   $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$.
- \end{lemma}
- \begin{proof}
-   \begin{enumproof}
-   \item $\exists r \in \var{localityReg}(g,W) \ldotp$ $\exists
-     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
-     \iota^\pwl_{\start',\addrend'}$: by definition of $\writeCond{}$.
-   \item Suffices: $W(r)\nsupsim[n-1] \iota^\nwl_{\start',\addrend'}$. By
-     definition of $\writeCond{}$
-   \item $W(r)\nsupsim[n-1] \iota^\pwl_{\start',\addrend'} \nsupsim[n-1]
-     \iota^\nwl_{\start',\addrend'}$: follows by Lemma~\ref{lem:nwl-subset-pwl}.
-   \end{enumproof}
- \end{proof}
-
  \begin{lemma}[Conditions for restrict instruction are sufficient]
@@
  \begin{proof}
-   By inspection of the definition of $\stdvr(W)$, everything follows trivially
-   except the result from lemma~\ref{lem:pwl-writecond-implies-nwl}.
-
-    If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$
-
-  \end{enumproof}
+   TODO
  \end{proof}

### 2017-02-03 local-capability-machine.tex

@@
 
+ \begin{lemma}[pwl writecond implies nwl]
+   \label{lem:pwl-writecond-implies-nwl}
+   If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
+   $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}$.
+ \end{lemma}
+ \begin{proof}
+   \begin{enumproof}
+   \item $\exists r \in \var{localityReg}(g,W) \ldotp$ $\exists
+     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
+     \iota^\pwl_{\start',\addrend'}$: by definition of $\writeCond{}$.
+   \item Suffices: $W(r)\nsupsim[n-1] \iota^\nwl_{\start',\addrend'}$. By
+     definition of $\writeCond{}$
+   \item $W(r)\nsupsim[n-1] \iota^\pwl_{\start',\addrend'} \nsupsim[n-1]
+     \iota^\nwl_{\start',\addrend'}$: follows by Lemma~\ref{lem:nwl-subset-pwl}.
+   \end{enumproof}
+ \end{proof}
+
+ \begin{lemma}[execCond implies entryCond]
+   \label{lem:execCond-implies-entryCond}
+   If $\npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W)$ then
+     $\npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)$.
+ \end{lemma}
+ \begin{proof}
+   \begin{enumproof}
+   \item Assume: $n' < n$, $W' \future W$ where $\gl = \local \Rightarrow
+     \future = \futurewk$ and $\gl = \glob \Rightarrow \future = \futurestr$
+     \\
+     Suffices: $\npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')$
+   \item Case $\addr \in [\start,\addrend]$: Follows from the definition of
+     $\execCond{}$.
+   \item Case $\addr \not\in[\start,\addrend]$: Follows by
+     Lemma~\ref{lem:failed-obs-stder}.
+   \end{enumproof}
+ \end{proof}
+
  \begin{lemma}[Conditions for restrict instruction are sufficient]
@@
  \begin{proof}
-   TODO
+   By inspection of the definition of $\stdvr(W)$, everything follows trivially
+   except the following.
+
+   \begin{enumproof}
+   \item If $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)$ then
+     $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)$: holds
+     by lemma~\ref{lem:pwl-writecond-implies-nwl}.
+   \item If $\npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W)$ then
+     $\npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)$.
+   \end{enumproof}
  \end{proof}

### 2017-02-03 local-capability-machine.tex
Indentation
changes/corrections to f1 proof

@@
     \step[k]
-                  (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}''}\update{r_1}{\dots}\update{r_l}{c_l},m'')
+    (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}''}\update{r_1}{\var{len}_{\var{ar}}}\update{r_l}{c_l},m'')
   \]
@@
   \begin{itemize}
-                \item $\cdots$ represents the needed size for the activation record. 
+  \item $\var{len}_{\var{ar}}$ is the needed size for the activation record. 
   \item \ref{f1:mallocsat}.\ and (downwards closure) gives us the needed memory segment satisfaction.
@@
   \[
-                  (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}''}\update{r_1}{\dots}\update{r_l}{c_l},m'')
+    (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}''}\update{r_1}{\var{len}_{\var{ar}}}\update{r_l}{c_l},m'')
     \step[j']
@@
   \item $\heapSat[\ms_{\malloc}'']{n-j-j'}{[0 \mapsto \iota_{\malloc}']}$ \label{f1:mallocsat}
-                \item $\dom(\hs_{\var{ar}}) = [b,e]$
+  \item $\dom(\hs_{\var{ar}}) = [b,e]$, and $e-b = \var{len}_{\var{ar}}$
   \item $c_l = ((\rwx,\glob),b,e,b)$ \label{test}
@@
   \item $\ms_{\var{ar}}'$ contains the activation record, i.e., $c_l$, $c_{f1}^{(3)}$ (the return address in f1), and activation code.
-                \item $c_{\var{ar}}' = ((\entry,\local)b,e,b+\olf)$ where $b+\olf$ is the first address of the activation code.
+  \item $c_{\var{ar}}' = ((\entry,\local)b,e,b+\var{offset})$ where $b+\var{offset}$ is the first address of the activation code.
   \end{itemize}
@@
     [1 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
-                  [2 \mapsto \iota_{f1}]
-                  [3 \mapsto \iota_{\var{ar}}]
-                  [4 \mapsto \iota_l]
-                  [5 \mapsto \iota_\flag]
-                  [5 \mapsto \iota_\link]$
-                \end{itemize}
-                \lau{Need to change this to use the new standard regions}
-                where
-                \begin{itemize}
-                \item $\iota_{f1}$, $\iota_{\var{ar}}$, $\iota_l$, $\iota_\flag$, and $\iota_\link$ are permanent islands that only accepts respectively $\ms_{f1}$, $\ms_{\var{ar}}$, $\ms_l$, $\ms_\flag$, and $\ms_\link$.
+    [2 \mapsto \iota^\sta (\perm,\ms_{f1} \uplus \ms_{\var{ar}} \uplus \ms_l \uplus \ms_\flag)]
+    [3 \mapsto \iota^{\sta,u}(\perm,\ms_\link)]$
   \end{itemize}
@@
     \begin{itemize}
-                  \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}
-                    \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: see Lemma~\ref{lem:nwlp-subset-pwl}.
+    \item Show: $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv} \nsubsim[n] \iota^\pwl_{\start_\adv,\addrend_\adv}$: Follows from Lemma~\ref{lem:nwlp-subset-pwl}.
     \end{itemize}
@@
   \end{enumerate}
-                Let $n' = n - j - j'$ and show
+  Let $n' = n - j - j'-k-k'$ and show
   \begin{enumproof}[resume]
-                \item $n' \leq n$
-                  \begin{enumproof}
-                  \item trivial
-                  \end{enumproof}
   \item $\heapSat[\ms]{n'}{W}$
@@
         \begin{enumproof}
-                      \item Show $\forall a \in [\start_\adv,\addrend_\adv] \ldotp \npair[n'-1]{\ms(a)} \in \stdvr(W)$
-                      \end{enumproof}
-                      (non-local requirement trivial - only capability if the link one, and it is $\glob$).
-                      \begin{enumproof}
-                      \item $a \neq \start_\adv$ : trivial, contains instruction only.
+        \item Show $\forall a \in [\start_\adv,\addrend_\adv] \ldotp (\npair[n'-1]{\ms(a)} \in \stdvr(W) \land \nonlocal{\ms(a)})$
+        \end{enumproof}
+        \begin{enumproof}
+        \item $a \neq \start_\adv$ : trivial, contains instruction only and they are non-local.
         \item $a = \start_\adv$: show $((\readonly,\glob),\link,\link+1,\link) \in \stdvr(W)$\\
-                        SFTS $\iota_{\link} \nsubsim[n'] \iota^\pwl_{\link,\link+1}$\\
-                        $\iota_{\link}$ accepts only $\ms_\link$, so show:
-                        \begin{enumproof}
-                        \item $\npair[n'-1]{c_\malloc} \in \stdvr(W)$: We assume
-                          this is indeed the case.\dominique{TODO: prove}
+          $\glob$ capabilities are non-local.\\
+          SFTS $\iota^{\sta,u}(\perm,\ms_\link) \nsubsim[n'] \iota^\pwl_{\link,\link+1}$ which follows from Lemma~\ref{lem:stau-subset-pwl}.
+        \item $\npair[n']{\ms_\link} \in H^{\sta,u}(1)(W)$:\\
+          This boils down to showing:
+          \begin{enumproof}
+          \item $\npair[n'-1]{c_\malloc} \in \stdvr(W)$: Follows from Lemma~\ref{lem:malloc-in-vr}.
+
           \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: for $n'' <
             n'-1$ and $W' \futurestr W$ show:
-                          \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stdvr(W')$.
-                          Use FTLR.\dominique{no: by induction hypothesis?} \label{f1:adv}
-                        \end{enumproof}
-                      \end{enumproof}
-                    \item The remaining islands are designed to accept their designated islands.
+            \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stder(W')$.
+            Use FTLR. \label{f1:adv}
+          \end{enumproof}
+        \end{enumproof}
+      \item The last case follows from Lemma~\ref{lem:mem-sat-static}
       \end{enumproof}
@@
       \begin{enumproof}
-                    \item Let $n'' < n'$ and $W' \futurewk W$ be given and show $\npair[n'']{\updatePcPerm{c_{\var{ar}}'}} \in \stdvr(W')$\\
-                      Let $n^{(3)} \leq n''$, $\heapSat[\ms']{n''}{W'}$, and $\npair[n^{(3)}]{\reg}$ be given\\
+      \item Let $n'' < n'$ and $W' \futurewk W$ be given and show $\npair[n'']{\updatePcPerm{c_{\var{ar}}'}} \in \stder(W')$\\
+        Let $n^{(3)} \leq n''$, $\heapSat[\ms']{n^{(3)}}{W'}$, and $\npair[n^{(3)}]{\reg}$ be given\\
         Show: $\npair[n^{(3)}]{(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms')} \in \observations(W')$\\
-                      Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms' \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k''$, $m'$ and $\ms_{\var{frame}}$.
-Due to $\heapSat[\ms']{n''}{W'}$, $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
+        Assume $(\reg\update{\pcreg}{\updatePcPerm{c_{\var{ar}}'}},\ms' \uplus \ms_{\var{frame}}) \step[k''] (\halted, m')$, for some $k'' \leq n^{(3)}$, $m'$ and $\ms_{\var{frame}}$.
+Due to $\heapSat[\ms']{n^{(3)}}{W'}$, $\ms_{f1}$, $\ms_\flag$, $\ms_{\var{ar}}'$, and $\ms_l$ are unchanged.
 \\
@@
 In other words, there were no changes to the memory.\\
-                      Use $W'$, $\ms_r = \emptyset$, and $\ms'$ to get the desired result, i.e., $m' = \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n''-k''}{W'}$ (using downwards closure of memory satisfaction).
-                    \end{enumproof}
-                  \item case: $\npair[n']{0} \in \stdvr(W)$
-                    \begin{enumproof}
-                    \item Trivial
-                    \end{enumproof}
+        Use $W'$, $\ms_r = \emptyset$, and $\ms'$ to get the desired result, i.e., $m' = \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n^{(3)}-k''}{W'}$ (using downwards closure of memory satisfaction).
+      \end{enumproof}
+    \item case: $\npair[n']{0} \in \stdvr(W)$ (the contents remaining registers)\\
+      Trivial to show.
     \end{enumproof}
@@
   \[
-                  \npair[n']{(\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_1}{c_{\var{adv}}}\update{r_0}{c_{ar}'},\ms)} \in \observations(W)
+    \npair[n']{(\reg_0\update{\pcreg}{\updatePcPerm{c_{\var{adv}}}}\update{r_1}{c_{\var{adv}}}\update{r_0}{c_{ar}'},m^{(3)})} \in \observations(W)
   \]
@@
   $W' \futurestr W$, $\ms_r$ and $\ms'$ s.t.
-$m' = \ms_r \uplus \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n' - \dots}{W'}$.
+$m' = \ms_r \uplus \ms' \uplus \ms_{\var{frame}}$ and $\heapSat[\ms']{n}{W'}$.
 As $\iota_\flag$ is a permanent region, we know it is still in $W'$, so $m'(\flag) = 0$.
@@
       \item $\ms_\malloc$: governed by $\iota_{\malloc,0}$, use malloc specification.
-      \item $\ms_\flag \uplus \ms_{f2}$: governed by island 1, only this memory segment is accepted.
-      \item $\ms_\link$: governed by island 2, only this memory segment is accepted.
+      \item $\ms_\flag \uplus \ms_{f2}$: governed by region 1, only this memory segment is accepted.
+      \item $\ms_\link$: governed by region 2, only this memory segment is accepted.
       \item $\ms_\adv$:\\

### 2017-02-03 local-capability-machine.tex
small update to f1 lemma
typo

@@
   \begin{itemize}
-  \item $c_\malloc$ satisfies the specification for malloc.
+  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
   \end{itemize}
@@
 If it was temporary, then it is there because it was just reinstated.
-If it was revoked in $W_3$, then it is still there because the only reinstated island were the temporary ones in $W_3$.
+If it was revoked in $W_3$, then it is still there because the only reinstated region were the temporary ones in $W_3$.
 

### 2017-02-03 local-capability-machine.tex
updated f1 to use fetch
Added new proof of f2s correctness using the scall lemma

@@
                   Clear all the register that are not $\pcreg$, $r$, $r_0$ or in $\bar{r}_{\var{args}}$.
-                \item [Jump] Jump to $r$
+                \item [Jump] Jump to register $r$
                 \item [Activation code] The activation code does the following:
@@
     store r_l 1
-    call adv([],[r_l])
+    fetch r_adv adv
+    call r_adv([],[r_l])
     assert r_l 1
@@
           SFTS $\iota^{\sta,u}(\perm,\ms_\link) \nsubsim[n'] \iota^\pwl_{\link,\link+1}$ which follows from Lemma~\ref{lem:stau-subset-pwl}.
+        \end{enumproof}
       \item $\npair[n']{\ms_\link} \in H^{\sta,u}(1)(W)$:\\
@@
         \end{enumproof}
-        \end{enumproof}
+
       \item The last case follows from Lemma~\ref{lem:mem-sat-static}
@@
     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
     \reg & \in \Regs \\
@@
   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-  \item $\hs_\adv$ is only instructions and possibly a global read-only capability for $\hs_\link$
+  \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
   \end{itemize}
@@
 \end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:correctness-f2} (using \texttt{scall} lemma)]
+    Let $n$ be given and make the assumptions of the lemma. If we can show
+    \begin{equation}
+      \label{pf:f2-sc:sfts-obs}
+      \npair{(\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},\ms \uplus \ms_\stk)} \in \observations(W)
+    \end{equation}
+    for 
+    \[
+      \ms = \hs_{f2} \uplus 
+            \hs_\flag \uplus                
+            \ms_{\var{link}} \uplus 
+            \hs_\adv \uplus 
+            \ms_{\malloc} 
+    \]
+    and
+    \[
+      W = [0 \mapsto \iota_{\malloc,0}]
+          [1 \mapsto \iota^\sta (\perma, \ms_{f2} \uplus \ms_\flag)]
+          [2 \mapsto \iota^{\sta,u} (\perma, \ms_\link)]
+          [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
+    \]
+    then we are done as we by assumption has
+    \[
+      (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m')
+    \]
+    so \ref{pf:f2-sc:sfts-obs} gives us a $W' \futurestr W$ where $W'$ satisfy part of $m'$. As $\ms_\flag$ is governed by a $\perma$ region, so it is unchanged. In other words
+    \[
+      m'(\flag) = 0
+    \]
+    So it suffices to show \ref{pf:f2-sc:sfts-obs}. To this end use Lemma~\ref{lem:anti-red-obs}. Let $\ms_f$ be given, then
+  \begin{multline*}
+    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},\ms \uplus \ms_\stk \uplus \ms_f) \step[k] \\
+    (\reg',\ms \uplus \ms_\stk' \uplus \ms_f)
+  \end{multline*}
+  where
+  \begin{itemize}
+  \item \lookingat{(\reg',\ms)}{\scall{r_\adv}{}{r_l}}{c_{\var{next}}}
+  \item $c_{\var{next}}$ is $c_{f2}$ that points to the instruction after the \texttt{scall}.
+  \item \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\unused}
+    \begin{itemize}
+    \item for some $\ms_\unused$ where $\ms_\stk' = [\start_\stk \mapsto 1] \uplus \ms_\unused$.
+    \end{itemize}
+  \item $\reg'(r_\adv) = c_\adv$
+  \end{itemize}
+  In order to show the observation part necessary for Lemma~\ref{lem:anti-red-obs}, we use the "\texttt{scall} works"-Lemma (Lemma~\ref{lem:scall-works}). Show the following
+  \begin{enumproof}
+  \item $\memSat{n-k}{\ms}{W}$\\
+    Use Lemma~\ref{lem:disj-mem-sat} with
+    \begin{enumproof}
+    \item $\memSat[n-k]{\ms_{f2} \uplus \ms_\flag}{[1 \mapsto \iota^\sta (\perma, \ms_{f2} \uplus \ms_\flag)]}$\\
+      Lemma~\ref{lem:mem-sat-static}
+      \item $\memSat[n-k]{\ms_\adv \uplus \ms_\malloc \uplus \ms_\link}{W_{\var{part}}}$ \\
+        where
+        \[
+          W_{\var{part}} = [0 \mapsto \iota_{\malloc,0}][2 \mapsto \iota^{\sta,u} (\perma, \ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]
+        \]
+        This amounts to
+        \begin{enumproof}
+        \item $\npair[n-k-1]{\ms_\malloc} \in H \; 1 \; W_{\var{part}}$ where $H$ is the interpretaion of the $\iota_{\malloc,0}$ region.\\
+          Follows from the malloc specification.
+        \item $\npair[n-k-1]{\ms_\adv} \in H^\nwl \; 1 \; W_{\var{part}}$\\
+          Can be shown using Lemma~\ref{lem:stau-subset-pwl}.
+        \item $\npair[n-k-1]{\ms_\link} \in H^{\sta,u} (\ms_\link) \; 1 \; W_{\var{part}}$\\
+          This amounts to showing
+          \begin{enumproof}
+            \item $\npair[n-k-2]{c_\malloc} \in \stdvr(W_{\var{part}})$
+              Follows from Lemma~\ref{lem:malloc-in-vr}.
+            \item $\npair[n-k-2]{c_\adv} \in \stdvr(W_{\var{part}})$\\
+              Follows from Theorem~\ref{thm:ftlr} using Lemma~\ref{lem:nwlp-subset-pwl}.
+          \end{enumproof}
+        \end{enumproof}
+    \end{enumproof}
+  \item Hyp-Callee\\
+    Assume
+    \begin{itemize}
+    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{act}} \uplus \ms_{\mathit{unused}}')$,
+    \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
+    \item $\memSat[n-k-1]{\ms''}{W'}$
+    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
+    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{c_\adv},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r_\adv \mapsto c_\adv]$ 
+    \item $\npair[n-k-1]{c_{\mathit{ret}}} \in \stdvr(W')$
+    \item $\npair[n-k-1]{c_\stk'} \in \stdvr(W')$
+    \end{itemize}
+    Show
+    \[
+      (n-k-1,(\reg',\ms'')) \in \observations(W')
+    \]
+    By Theorem~\ref{thm:ftlr} we get
+    \[
+      \npair[n-k-1]{\updatePcPerm{c_\adv}} \in \stder(W')
+    \]
+    getting the desired result amounts to\footnote{We have memory satisfaction by assumption and the above entails the register-file is in the register-file relation.}
+    \begin{enumproof}
+      \item $\npair[n-k-1]{c_\adv} \in \stdvr{W}$ \\
+        To this end let $n' < n-k-1$ and $W'' \futurestr W'$ be given and show
+        \[
+          \npair[n']{\updatePcPerm{c_\adv}} \in \stder(W'')
+        \]
+        Follows from Theorem~\ref{thm:ftlr} and Lemma~\ref{lem:nwlp-subset-pwl}.
+    \end{enumproof}
+  \item Hyp-Cont\\
+    Assume
+    \begin{itemize}
+    \item $n' \leq n-2$
+    \item $W'' \futurewk \revokeTemp{W}$
+    \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
+    \item $\reg''(\pcreg) = c_{\mathit{next}}$
+    \item $\reg'' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
+    \end{itemize}
+    and show
+    \[
+      \npair[n']{(\reg'',\ms'' \uplus [\start_\stk \mapsto 1] \uplus \ms_{\var{unused}}'')} \in \observations(W'')
+    \]
+    From $\memSat[n']{\ms''}{\revokeTemp{W''}}$, we get that $\ms_{f2}$ is unchanged. Given a frame $\ms_f'$ and assuming $n'$ is sufficiently large, the execution continues as follows:
+    \[
+      (\reg'',\ms'' \uplus [\start_\stk \mapsto 1] \uplus \ms_{\var{unused}}'' \uplus \ms_f) \step[k] (\halted,\ms'' \uplus [\start_\stk \mapsto 1] \uplus \ms_{\var{unused}}'' \uplus \ms_f)
+    \]
+    because 1 is popped of the stack to a register, then it is compared with 1 in the assertion, so the assertion succeeds and halts immediately after.
+
+    By assumption we had $\memSat[n']{\ms''}{\revokeTemp{W''}}$ which gives us exactly the memory satisfaction required by $\observations(W'')$.
+  \end{enumproof}
+\end{proof}
+
 \begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
@@
       \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg(r)]$ 
+      \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r \mapsto \reg(r)]$ 
     \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
@@
 
-    then we have that $(n-1,(\reg',\ms'')) \in \observations(W')$
+    then we have that $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
   \item[Hyp-Cont] If
@@
 
-    then we have that $(n', (\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')) \in \observations(W'')$
+    then we have that $\npair[n']{(\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')} \in \observations(W'')$
   \end{itemize}

### 2017-02-04 csf-2017/paper.tex
csf-2017 directory and files added

\documentclass[compsoc,conference,letterpaper,fleqn]{IEEEtran}
\usepackage{hyperref}

\usepackage{cite}
\usepackage{mathpartir}
\usepackage{url}
\usepackage{comment}
\usepackage{amsmath}

\begin{document}

\title{Reasoning about a Capability Machine with Local Capabilities\\
 Provably Safe Stack and Return Pointer Management (without OS Support).}

\author{%
\IEEEauthorblockN{Lau Skorstengaard}
\IEEEauthorblockA{Aarhus University, Denmark\\
Email: lau@cs.au.dk} \and
\IEEEauthorblockN{Dominique~Devriese}
  \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
Email: dominique.devriese@cs.kuleuven.be} \and
\IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
Email: birkedal@cs.au.dk}}

\maketitle

\begin{abstract}
  abc
\end{abstract}

\section{Introduction}
\label{sec:introduction}

\begin{comment}

*** Introduce capability machines
*** Enforcing properties on a capability machine

    Privacy of local state using memory capabilities.
    Encapsulation of private state using enter pointers or CHERI's ccall.
    (mention in passing secure compilation, full abstraction)

*** Well-bracketed control flow and encapsulation of stack frames

    Hard to enforce using standard capabilities:
    - Well-bracketed control flow can be broken if the adversary stores return pointers.
      (alternative: trusted stack)
    - Encapsulation of local stack frames can be broken if the adversary stores stack pointers.

*** Local capabilities

    - Introduced by CHERI for this purpose.
    - Explain how they can be used for passing return pointers and stack pointers.
    - But this is not watertight.
    - Example attack:
      + In a first invocation, adversary fills the entire unused part of the stack with copies of the current stack pointer
      + In a second invocation, adversary can load the old stack pointer of the stack
      + He now has access to private state that was pushed on the stack in between.
    - So:
      + Can we make it (provably) watertight?
      + Is the trusted stack necessary? Why not just pass return pointers locally?
      + General question: How to reason about capability machines and
        local capabilities?

*** Contributions

    - Formal model of a simple but representative capability machine with local capabilities
    - Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
      + untrusted adversary
      + higher-order code: callbacks passed to and received from the adversary
      + efficient stack management
      + no OS support
    - Logical relation for reasoning about code in this capability machine
    - Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
    - Technical contributions: reuse existing ideas in a new way:
      + replace biorthogonal closure by a single orthogonal closure
        because assembly languages remove the distinction between the
        continuation and the arguments
	LB: we need to relate this carefully to Hur-Dreyer, who used
        biorthogonality even though they also worked with an assembly
        language. (If I recall correctly, they assumed some properties
        of the low level language, which is why the biortho was the
        right ting, but we need to check.)
	LS: It is my impression that they were interested in the relation
	between high-level and low-level programs. They are therefore in
	particular interested in low-level programs compiled from high-level
	programs which means that the continuation is always invoked in the 
	same way. Our realisation was that anything we pass as an argument 
	in a "call" can be used as a continuation, so the continuation 
	relation was redundant. Eventhough we do try to give our programs
	some structure using the macros, our logical relation is strong enough
	to handle unstructured programs as well.
      + STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
    - Demonstrate all of this on several examples:
      + security examples
      + the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
      + some compartmentalisation result?
      + whatever else we do..

*** Outline

** A capability machine with local capabilities

   Formalisation

   Explain our model of linking

** Stack and return pointer management using local capabilities

   - Issues to consider:
     + no global store-local capabilities
     + clearing the stack
     + clearing (some) registers
     + constructing return pointers
     + constructing callbacks
     + invoking the adversary's callbacks
   - Reusable macro-instructions for stack and return pointer management
     + taking care of the stuff above

** LB: maybe we can present the examples informally before the logical relation and then return to them after the logical relation (to present the formal properties) ?

** Logical relation

   - Intuition
   - Worlds
   - Logical relation
   - Fundamental theorem
   - Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer

** Examples
   - Ticket dispenser
   - The awkward example and variants
   - A sandboxing example?

     For example, an untrusted advertisement scenario with initialization code
     that registers a redraw callback. The redraw callback gets temporary
     read-write access to a framebuffer.

   - Some compartmentalisation result?

** Discussion, future work
   - A general well-bracketed control flow result?
     + what would that result say?
     + possible idea: fully abstract compilation from an assembly language with
       a trusted stack to one without
     + the LR and some of the lemmas already imply well-bracketed control flow, as seen in examples
   - Relation to local parameters in Scala, Algol/Pascal second-class function parameters?
   - Stack clearing realistic?
   - Non-modularity of heap allocation requirement for adversary callbacks

** Related work

   - Dreyer-Neis-Birkedal
   - CHERI papers
   - Akram's thesis
   - http://2016.splashcon.org/event/splash-2016-oopsla-gentrification-gone-too-far-affordable-2nd-class-values-for-fun-and-co-effect
   - other papers that enforce well-bracketed control flow at low level
     (using a trusted stack manager)
     + http://ieeexplore.ieee.org/abstract/document/7536364/
     + http://ieeexplore.ieee.org/abstract/document/7536366/
     + other stuff?

** Conclusion

\end{comment}

\section*{Acknowledgements}
\label{sec:acknowledgements}

This research was supported in part by the ModuRes Sapere Aude Advanced Grant from The Danish Council for Independent Research for the Natural Sciences (FNU).

\end{document}

### 2017-02-04 local-capability-machine.tex
put the old f2 proof in a comment (not sure whether the proof still holds after our changes)
existence proof typo fixed

@@
 
+\begin{comment}
 \begin{proof}[Proof of lemma~\ref{lem:correctness-f2}]
@@
 \end{proof}
-
+\end{comment}
 \lau{I have skipped the following program in favour of the one in the next section. }
@@
   The theorem follows from a more general solution theorem for the
-  category of $P$ preordered \cofes, see~\cite{bizjak-note}. 
+  category of $P$ preordered \cofes,
+  see~\cite{Birkedal:2010:TCS:411:4102-4122} and \cite{bizjak-note}. 
   We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to
@@
                                             \UPred{\HeapSegments})))),
-                                        \futurestr)
+                                        \futurewk)
   \end{align*}

### 2017-02-04 csf-2017/paper.tex
Update the paper template using the org mode Latex export.

@@
 
-\begin{comment}
-
-*** Introduce capability machines
-*** Enforcing properties on a capability machine
+\subsection{Introduce capability machines}
+\subsection{Enforcing properties on a capability machine}
 
@@
 
-*** Well-bracketed control flow and encapsulation of stack frames
+\subsection{Well-bracketed control flow and encapsulation of stack frames}
 
 Hard to enforce using standard capabilities:
-    - Well-bracketed control flow can be broken if the adversary stores return pointers.
+\begin{itemize}
+\item Well-bracketed control flow can be broken if the adversary stores return pointers.
 (alternative: trusted stack)
-    - Encapsulation of local stack frames can be broken if the adversary stores stack pointers.
+\item Encapsulation of local stack frames can be broken if the adversary stores stack pointers.
+\end{itemize}
 
-*** Local capabilities
+\subsection{Local capabilities}
 
-    - Introduced by CHERI for this purpose.
-    - Explain how they can be used for passing return pointers and stack pointers.
-    - But this is not watertight.
-    - Example attack:
-      + In a first invocation, adversary fills the entire unused part of the stack with copies of the current stack pointer
-      + In a second invocation, adversary can load the old stack pointer of the stack
-      + He now has access to private state that was pushed on the stack in between.
-    - So:
-      + Can we make it (provably) watertight?
-      + Is the trusted stack necessary? Why not just pass return pointers locally?
-      + General question: How to reason about capability machines and
+\begin{itemize}
+\item Introduced by CHERI for this purpose.
+\item Explain how they can be used for passing return pointers and stack pointers.
+\item But this is not watertight.
+\item Example attack:
+\begin{itemize}
+\item In a first invocation, adversary fills the entire unused part of the stack with copies of the current stack pointer
+\item In a second invocation, adversary can load the old stack pointer of the stack
+\item He now has access to private state that was pushed on the stack in between.
+\end{itemize}
+\item So:
+\begin{itemize}
+\item Can we make it (provably) watertight?
+\item Is the trusted stack necessary? Why not just pass return pointers locally?
+\item General question: How to reason about capability machines and
 local capabilities?
+\end{itemize}
+\end{itemize}
 
-*** Contributions
+\subsection{Contributions}
 
-    - Formal model of a simple but representative capability machine with local capabilities
-    - Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
-      + untrusted adversary
-      + higher-order code: callbacks passed to and received from the adversary
-      + efficient stack management
-      + no OS support
-    - Logical relation for reasoning about code in this capability machine
-    - Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
-    - Technical contributions: reuse existing ideas in a new way:
-      + replace biorthogonal closure by a single orthogonal closure
+\begin{itemize}
+\item Formal model of a simple but representative capability machine with local capabilities
+\item Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
+\begin{itemize}
+\item untrusted adversary
+\item higher-order code: callbacks passed to and received from the adversary
+\item efficient stack management
+\item no OS support
+\end{itemize}
+\item Logical relation for reasoning about code in this capability machine
+\item Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
+\item Technical contributions: reuse existing ideas in a new way:
+\begin{itemize}
+\item replace biorthogonal closure by a single orthogonal closure
 because assembly languages remove the distinction between the
 continuation and the arguments
+\end{itemize}
+\end{itemize}
 LB: we need to relate this carefully to Hur-Dreyer, who used
@@
 to handle unstructured programs as well.
-      + STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
-    - Demonstrate all of this on several examples:
-      + security examples
-      + the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
-      + some compartmentalisation result?
-      + whatever else we do..
+\begin{itemize}
+\item STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
+\end{itemize}
+\begin{itemize}
+\item Demonstrate all of this on several examples:
+\begin{itemize}
+\item security examples
+\item the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
+\item some compartmentalisation result?
+\item whatever else we do..
+\end{itemize}
+\end{itemize}
 
-*** Outline
+\subsection{Outline}
 
-** A capability machine with local capabilities
+\section{A capability machine with local capabilities}
 
@@
 
-** Stack and return pointer management using local capabilities
+\section{Stack and return pointer management using local capabilities}
 
-   - Issues to consider:
-     + no global store-local capabilities
-     + clearing the stack
-     + clearing (some) registers
-     + constructing return pointers
-     + constructing callbacks
-     + invoking the adversary's callbacks
-   - Reusable macro-instructions for stack and return pointer management
-     + taking care of the stuff above
+\begin{itemize}
+\item Issues to consider:
+\begin{itemize}
+\item no global store-local capabilities
+\item clearing the stack
+\item clearing (some) registers
+\item constructing return pointers
+\item constructing callbacks
+\item invoking the adversary's callbacks
+\end{itemize}
+\item Reusable macro-instructions for stack and return pointer management
+\begin{itemize}
+\item taking care of the stuff above
+\end{itemize}
+\end{itemize}
 
-** LB: maybe we can present the examples informally before the logical relation and then return to them after the logical relation (to present the formal properties) ?
+LB: maybe we can present the examples informally before the logical relation and then return to them after the logical relation (to present the formal properties) ?
 
-** Logical relation
+\section{Logical relation}
 
-   - Intuition
-   - Worlds
-   - Logical relation
-   - Fundamental theorem
-   - Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer
+\begin{itemize}
+\item Intuition
+\item Worlds
+\item Logical relation
+\item Fundamental theorem
+\item Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer
+\end{itemize}
 
-** Examples
-   - Ticket dispenser
-   - The awkward example and variants
-   - A sandboxing example?
+\section{Examples}
+\begin{itemize}
+\item Ticket dispenser
+\item The awkward example and variants
+\item A sandboxing example?
 
@@
 
-   - Some compartmentalisation result?
+\item Some compartmentalisation result?
+\end{itemize}
 
-** Discussion, future work
-   - A general well-bracketed control flow result?
-     + what would that result say?
-     + possible idea: fully abstract compilation from an assembly language with
+\section{Discussion, future work}
+\begin{itemize}
+\item A general well-bracketed control flow result?
+\begin{itemize}
+\item what would that result say?
+\item possible idea: fully abstract compilation from an assembly language with
 a trusted stack to one without
-     + the LR and some of the lemmas already imply well-bracketed control flow, as seen in examples
-   - Relation to local parameters in Scala, Algol/Pascal second-class function parameters?
-   - Stack clearing realistic?
-   - Non-modularity of heap allocation requirement for adversary callbacks
+\item the LR and some of the lemmas already imply well-bracketed control flow, as seen in examples
+\end{itemize}
+\item Relation to local parameters in Scala, Algol/Pascal second-class function parameters?
+\item Stack clearing realistic?
+\item Non-modularity of heap allocation requirement for adversary callbacks
+\end{itemize}
 
-** Related work
+\section{Related work}
 
-   - Dreyer-Neis-Birkedal
-   - CHERI papers
-   - Akram's thesis
-   - http://2016.splashcon.org/event/splash-2016-oopsla-gentrification-gone-too-far-affordable-2nd-class-values-for-fun-and-co-effect
-   - other papers that enforce well-bracketed control flow at low level
+\begin{itemize}
+\item Dreyer-Neis-Birkedal
+\item CHERI papers
+\item Akram's thesis
+\item \url{http://2016.splashcon.org/event/splash-2016-oopsla-gentrification-gone-too-far-affordable-2nd-class-values-for-fun-and-co-effect}
+\item other papers that enforce well-bracketed control flow at low level
 (using a trusted stack manager)
-     + http://ieeexplore.ieee.org/abstract/document/7536364/
-     + http://ieeexplore.ieee.org/abstract/document/7536366/
-     + other stuff?
+\begin{itemize}
+\item \url{http://ieeexplore.ieee.org/abstract/document/7536364/}
+\item \url{http://ieeexplore.ieee.org/abstract/document/7536366/}
+\item other stuff?
+\end{itemize}
+\end{itemize}
 
-** Conclusion
-
-\end{comment}
+\section{Conclusion}
 
@@
 This research was supported in part by the ModuRes Sapere Aude Advanced Grant from The Danish Council for Independent Research for the Natural Sciences (FNU).
+Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
 

### 2017-02-04 csf-2017/paper.tex
added part of intro

@@
 
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+
 \begin{document}
@@
 \author{%
-\IEEEauthorblockN{Lau Skorstengaard}
+\IEEEauthorblockN{Lau~Skorstengaard}
 \IEEEauthorblockA{Aarhus University, Denmark\\
@@
 \label{sec:introduction}
+%%% a)
+% Current low-level protection coarse grained memory protection or
+% properties of high-level languages are not really enforced?
+
+% Capability machines offer fine-grained memory protection
+A capability machine is a low-level machine that offers fine-grained
+memory protection.
+% Capability machines and unforgeable tokens for memory access. 
+This is done by introducing unforgeable tokens to
+the machine. These tokens are called capabillities and grant a kind of
+authority and a range of authority. We will refer to the capabilities
+that grant a combination of read, write, or execute permission as memory
+capabilities. Capabilities with the necessary permissions must be
+presented every time an instruction that manipulates the memory is
+executed.
+
+Memory capabilities provide memory protection, but they do not provide
+a way to set up security domains. When you jump to an untrusted piece
+of code that is supposed to jump back to you, then you have to either
+revoke all your capabilities or pass to the piece of code you do not
+trust. This short coming can be overcome by adding something like the
+enter capability from the M-Machine\todo{add reference}. The enter
+capability is an opaque capability which can only be used for a
+jump. When jumped to it grants read and execute permission which
+allows the code that is now executing to read capabilities stored in
+memory. 
+The CHERI processor's ccall achieves a similar property\todo{Add reference}.
+
+Capabilities are irrevocable, so when we pass a capability has to
+an untrusted program, then we have to assume that the program keeps it
+around indefenitely. This means that we cannot reuse the piece of
+memory that the capability governs essentially creating a memory
+leak. The CHERI processor has a special kind of capabilities called
+\emph{local capabilities}\todo{reference?} that introduces a simple
+kind of temporal information control which allows for a simple kind of
+capability revocation. This is achieved by adding a tag to every
+capability that marks it as either local or global. Local capabilities
+can only be written through a capability with a new permission called
+\emph{permit write local}.
+% More about local capabilities.
+
+With just memory capabilities, enter capabailities and local
+capabilities, the capability machine becomes very expressive and can
+enforce properties that high-level languages promisse. The memory
+capabilities enforce the privacy of local state as it is impossible
+for anyone to access your local state without the capability to do
+so. When this is combined with the enter capability,
+
+%%% b)
+
+% Examples of capability machines?
+
+% Enter capabilities
+
+% Local capabilities
+
+% High-level languages often provide guarantees such as encapsulation of private state and well-bracketedness of calls - often not clear how well this is enforced.
+Just enforced when interacting with other programs written in the same high-level language or is it also guaranteed when interacting with assembly programs.
+
+% High-level programs compiled to assembly can ensure these properties.
+
+%
 
@@
 \end{document}
+\grid

### 2017-02-05 csf-2017/paper.tex
note added

--- 2017-02-04 csf-2017/paper.tex
@@
 \end{itemize}
+LB: I think we should have some discussion of the flexibility /
+  strength of our LR.  The LR defines which computations we think of
+  as well-behaved.   
+ We should give simple / trivial examples of
+  code not in the LR. But we may also want to emphasize that the LR is 
+  not too restricting, e.g., it does not enforce a certain calling
+  scheme (examplied by Example \verb!f1! which
+  does not use the stack and the following examples which do use the
+  stack). This ties in to the discussion of biorthogonality, see the
+  following paragraphs.
+
 LB: we need to relate this carefully to Hur-Dreyer, who used
@@
       right ting, but we need to check.)
+
 LS: It is my impression that they were interested in the relation
@@
 in a "call" can be used as a continuation, so the continuation 
-relation was redundant. Eventhough we do try to give our programs
+relation was redundant. Even though we do try to give our programs
 some structure using the macros, our logical relation is strong enough

### 2017-02-05 csf-2017/paper.tex
minor notes
worlds

@@
 \documentclass[compsoc,conference,letterpaper,fleqn]{IEEEtran}
-\usepackage{hyperref}
 
 \usepackage{cite}
+\usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{thmtools}
+\usepackage{stmaryrd}
+\usepackage{natbib}
+\usepackage{url}
+\usepackage{array}
+\usepackage{arydshln}
+\usepackage{ifthen}
+\usepackage{ifpdf}
+\usepackage{verbatim}
 \usepackage{mathpartir}
-\usepackage{url}
-\usepackage{comment}
-\usepackage{amsmath}
+\usepackage{listings}
+\usepackage{hyperref}
+\lstset{
+  basicstyle=\ttfamily,
+  columns=fullflexible,
+  keepspaces=true,
+  mathescape
+}
+
+\usepackage{rotating}
+
+\usepackage{mathtools}
+\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
+\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
+
+\usepackage{enumitem}
+\newlist{enumproof}{enumerate}{10}
+\setlist[enumproof]{label*=\arabic*.}
+
+\usepackage{tikz}
+\usepackage{multirow,bigdelim}
+
+% %make latex preview-mode work with natbib...
+% \usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
+
+\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
+\declaretheorem[numbered=yes,name=Theorem,qed=$\blacksquare$]{theorem}
+\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
+\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}
+
+\newcommand{\forcenewline}{$\phantom{v}$\\}
+\newcommand{\judgment}[2]{\paragraph{#1}\hspace{\stretch{1}}\fbox{$#2$}}
+
+\newcommand{\update}[2]{[#1 \mapsto #2]}
+\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
+
+% Math notation
+\newcommand{\restrictfun}[1]{|_{#1}}
+\newcommand{\parfun}{\rightharpoonup}
+\newcommand{\finparfun}{\xrightharpoonup{\textit{\tiny{fin}}}}
+\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
+\newcommand{\monfun}{\xrightarrow{\textit{\tiny{mon}}}}
+\newcommand{\nefun}{\xrightarrow{\textit{\tiny{ne}}}}
+\newcommand{\fun}{\rightarrow}
+\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
+\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
+\renewcommand{\nsim}[1][n]{\stackrel{\tiny{#1}}{\simeq}}
+
+\newcommand\subsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\subset$}\cr
+      \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
+\newcommand\supsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\supset$}\cr
+      \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
+\newcommand{\nsubsim}[1][n]{\stackrel{\tiny{#1}}{\subsetsim}}
+\newcommand{\nsupsim}[1][n]{\stackrel{\tiny{#1}}{\supsetsim}}
+
+\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
+\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
+
+\newcommand{\union}{\mathbin{\cup}}
+\DeclareMathOperator{\dom}{dom}
+\newcommand{\blater}{\mathop{\blacktriangleright}}
+\newcommand{\id}{\var{id}}
+\newcommand{\undefined}{\mathit{undefined}}
+
+\newcommand{\powerset}[1]{\mathcal{P}(#1)}
+
+\newcommand{\false}{\mathit{false}}
+\newcommand{\true}{\mathit{true}}
+
+% cofes
+\newcommand{\cofe}{c.o.f.e.}
+\newcommand{\cofes}{\cofe{}'s}
+\newcommand{\CatC}{\mathbb{C}}
+\newcommand{\CatP}{\mathbb{P}}
+
+% Comments
+\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
+\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
+\newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+
+% Variables
+\newcommand{\var}[1]{\mathit{#1}}
+\newcommand{\hs}{\var{ms}}
+\newcommand{\ms}{\hs}
+\newcommand{\hv}{\var{hv}}
+\newcommand{\rv}{\var{rv}}
+\newcommand{\lv}{\var{lv}}
+\newcommand{\gl}{\var{g}}
+\newcommand{\pc}{\mathit{pc}}
+\newcommand{\pcreg}{\mathrm{pc}}
+\newcommand{\addr}{\var{a}}
+\newcommand{\offset}{\var{offset}}
+\newcommand{\word}{\var{w}}
+\newcommand{\start}{\var{base}}
+\newcommand{\addrend}{\var{end}}
+\newcommand{\pwlv}{\var{pwl}}
+\newcommand{\mem}{\var{mem}}
+\newcommand{\reg}{\var{reg}}
+\newcommand{\heapseg}{\var{ms}}
+\newcommand{\heap}{\var{mem}}
+\newcommand{\mode}{\var{mode}}
+\newcommand{\perm}{\var{perm}}
+\newcommand{\permp}{\var{permPair}}
+\newcommand{\roll}{\var{roll}}
+\newcommand{\instr}{\var{instr}}
+\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
+\newcommand{\adv}{\var{adv}}
+\newcommand{\msframe}{ms_\var{frame}}
+\newcommand{\link}{\var{link}}
+\newcommand{\stk}{\var{stk}}
+\newcommand{\flag}{\var{flag}}
+\newcommand{\nwl}{\var{nwl}}
+\newcommand{\pwl}{\var{pwl}}
+\newcommand{\sta}{\var{sta}}
+\newcommand{\cnst}{\var{cnst}}
+\newcommand{\olf}{\var{offsetLinkFlag}}
+\newcommand{\prp}{\var{prp}}
+\newcommand{\env}{\var{env}}
+\newcommand{\cls}{\var{cls}}
+\newcommand{\unused}{\var{unused}}
+\newcommand{\act}{\var{act}}
+
+% Memory projections
+\newcommand{\plainproj}[1]{\mathrm{#1}}
+\newcommand{\memheap}[1][\Phi]{#1.\plainproj{mem}}
+\newcommand{\memreg}[1][\Phi]{#1.\plainproj{reg}}
+
+\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{mem}.#2}{#3}}
+\newcommand{\updateReg}[3][\Phi]{#1\update{\plainproj{reg}.#2}{#3}}
+
+% Configuration end states
+\newcommand{\failed}{\textsl{failed}}
+\newcommand{\halted}{\textsl{halted}}
+
+% Functions
+\newcommand{\plainfun}[2]{
+  \ifthenelse{\equal{#2}{}}
+  {\mathit{#1}}
+  {\mathit{#1}(#2)}
+}
+\newcommand{\decode}{\plainfun{decode}{}}
+\newcommand{\encode}{\plainfun{encode}{}}
+\newcommand{\encodePerm}{\mathit{encodePerm}}
+\newcommand{\encodePermPair}{\plainfun{encodePermPair}{}}
+\newcommand{\encodeLoc}{\mathit{encodeLoc}{}}
+\newcommand{\decodePermPair}{\plainfun{decodePermPair}}
+\newcommand{\decodePerm}[1]{\plainfun{decodePerm}{#1}}
+\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}
+
+\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}{#1}}
+\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
+\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
+\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
+\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
+\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc}{#1}}
+
+\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
+\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
+\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
+\newcommand{\entryCond}[1]{\plainfun{enterCondition}{#1}}
+
+\newcommand{\revokeTemp}[1]{\plainfun{revokeTemp}{#1}}
+\newcommand{\erase}[2]{\floor*{#1}_{\{#2\}}}
+\newcommand{\activeReg}[1]{\plainfun{active}{#1}}
+
+% World operations
+\newcommand{\future}{\mathbin{\sqsupseteq}}
+\newcommand{\pub}{\var{pub}}
+\newcommand{\priv}{\var{priv}}
+\newcommand{\futurewk}{\mathbin{\sqsupseteq}^{\var{pub}}}
+\newcommand{\futurestr}{\mathbin{\sqsupseteq}^{\var{priv}}}
+\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
+\newcommand{\memSat}[3][n]{\heapSat[#2]{#1}{#3}}
+\newcommand{\memSatPar}[4][n]{\heapSat[#2]{#1 , #4}{#3}}
+
+\newcommand{\monwknefun}{\xrightarrow[\text{\tiny{$\futurewk$}}]{\textit{\tiny{mon, ne}}}}
+\newcommand{\monstrnefun}{\xrightarrow[\text{\tiny{$\futurestr$}}]{\textit{\tiny{mon, ne}}}}
+
+% Assembly labels
+\newcommand{\codelabel}[1]{\mathit{#1}}
+\newcommand{\init}{\codelabel{init}}
+\newcommand{\malloc}{\codelabel{malloc}}
+\newcommand{\counter}{\codelabel{counter}}
+\newcommand{\iocap}{\codelabel{iocap}}
+
+% Type(s)
+\newcommand{\type}[1]{\mathrm{#1}}
+\newcommand{\asmType}{\plaindom{AsmType}}
+
+% Domains
+\newcommand{\plaindom}[1]{\mathrm{#1}}
+\newcommand{\Caps}{\plaindom{Cap}}
+\newcommand{\Words}{\plaindom{Word}}
+\newcommand{\Addrs}{\plaindom{Addr}}
+\newcommand{\ExecConfs}{\plaindom{ExecConf}}
+\newcommand{\RegName}{\plaindom{RegisterName}}
+\newcommand{\Regs}{\plaindom{Reg}}
+\newcommand{\Heaps}{\plaindom{Mem}}
+\newcommand{\Mems}{\Heaps}
+%\newcommand{\HeapSegments}{\plaindom{MemSegment}}
+\newcommand{\HeapSegments}{\plaindom{MemSeg}}
+\newcommand{\MemSegments}{\HeapSegments}
+\newcommand{\Confs}{\plaindom{Conf}}
+\newcommand{\Instrs}{\plaindom{Instructions}}
+\newcommand{\nats}{\mathbb{N}}
+\newcommand{\ints}{\mathbb{Z}}
+\newcommand{\Perms}{\plaindom{Perm}}
+\newcommand{\Globals}{\plaindom{Global}}
+
+\newcommand{\Rel}{\plaindom{Rel}}
+\newcommand{\Rels}{\plaindom{Rels}}
+\newcommand{\States}{\plaindom{State}}
+\newcommand{\RegionNames}{\plaindom{RegionName}}
+\newcommand{\Regions}{\plaindom{Region}}
+\newcommand{\Reg}{\plaindom{Reg}}
+\newcommand{\Worlds}{\plaindom{World}}
+\newcommand{\Wor}{\plaindom{Wor}}
+\newcommand{\Worwk}{\Wor_{\futurewk}}
+\newcommand{\Worstr}{\Wor_{\futurestr}}
+\newcommand{\xiwk}{\xi_{\var{wk}}}
+\newcommand{\xistr}{\xi_{\var{str}}}
+\newcommand{\StorePred}{\plaindom{MemSegPred}}
+\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
+\newcommand{\DCPred}[1]{\plaindom{P}^\downarrow(#1)}
+
+\newcommand{\Views}{\plaindom{View}}
+
+% LR
+\newcommand{\intr}[2]{\mathcal{#1}}
+\newcommand{\valueintr}[1]{\intr{V}{#1}}
+\newcommand{\exprintr}[1]{\intr{E}{#1}}
+\newcommand{\contintr}[1]{\intr{K}{#1}}
+\newcommand{\regintr}[1]{\intr{R}{#1}}
+\newcommand{\stdvr}{\valueintr{\asmType}}
+\newcommand{\stder}{\exprintr{\asmType}}
+\newcommand{\stdrr}{\regintr{\asmType}}
+\newcommand{\stdkr}{\contintr{\asmType}}
+\newcommand{\observations}{\mathcal{O}}
+\newcommand{\npair}[2][n]{\left(#1,#2 \right)}
+
+% Reference register/memory
+\newcommand{\refreg}[1]{\lfloor #1 \rfloor}
+\newcommand{\refheap}[1]{\langle #1 \rangle_m}
+
+% Instructions
+% No arguments
+\newcommand{\zinstr}[1]{\mathtt{#1}}
+\newcommand{\fail}{\zinstr{fail}}
+\newcommand{\halt}{\zinstr{halt}}
+% One argument
+\newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
+\newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
+% Two arguments
+\newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
+\newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
+\newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
+\newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
+\newcommand{\geta}[2]{\twoinstr{geta}{#1}{#2}}
+\newcommand{\getb}[2]{\twoinstr{getb}{#1}{#2}}
+\newcommand{\gete}[2]{\twoinstr{gete}{#1}{#2}}
+\newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
+\newcommand{\getl}[2]{\twoinstr{getl}{#1}{#2}}
+\newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
+\newcommand{\store}[2]{\twoinstr{store}{#1}{#2}}
+\newcommand{\load}[2]{\twoinstr{load}{#1}{#2}}
+\newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
+% Three arguments
+\newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
+\newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
+\newcommand{\subseg}[3]{\threeinstr{subseg}{#1}{#2}{#3}}
+\newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}
+
+% Permissions
+\newcommand{\plainperm}[1]{\mathrm{#1}}
+\newcommand{\noperm}{\plainperm{o}}
+\newcommand{\readonly}{\plainperm{ro}}
+\newcommand{\readwrite}{\plainperm{rw}}
+\newcommand{\exec}{\plainperm{rx}}
+\newcommand{\entry}{\plainperm{e}}
+\newcommand{\rwx}{\plainperm{rwx}}
+% PWL permissions
+\newcommand{\readwritel}{\plainperm{rwl}}
+\newcommand{\rwl}{\readwritel}
+\newcommand{\rwlx}{\plainperm{rwlx}}
+
+% Global/local
+\newcommand{\local}{\plainperm{local}}
+\newcommand{\glob}{\plainperm{global}}
+
+\newcommand{\localityReg}{\var{localityReg}}
+\newcommand{\localReg}{\var{localReg}}
+\newcommand{\globalReg}{\var{globalReg}}
+
+% Views
+\newcommand{\plainview}[1]{\mathrm{#1}}
+\newcommand{\perma}{\plainview{perm}}
+\newcommand{\temp}{\plainview{temp}}
+\newcommand{\revoked}{\plainview{revoked}}
+
+% OP sem
+\newcommand{\diverge}[1][n]{\not\Downarrow_{#1}}
+\newcommand{\step}[1][]{\rightarrow_{#1}}
+
+% Conv defs
+\newcommand{\lookingat}[3]{\ensuremath{#1} \text{ is looking at } \ensuremath{#2} \text{ followed by } \ensuremath{#3}}
+\newcommand{\pointstostack}[3]{\ensuremath{#1} \text{ points to stack with } \ensuremath{#2} \text{ used and } \ensuremath{#3} \text{ unused}}
+\newcommand{\nonlocal}[1]{\ensuremath{#1} \text{ is non-local}}
+
+% Macros
+\newcommand{\scall}[3]{\mathtt{scall} \; #1([#2],[#3])}
 
@@
 \end{abstract}
+
+\section*{Terminology ?}
+
+\emph{capability safe}: 
+Our unary logical relation is a semantic \emph{definition}
+of which values for the
+pc / register-file / word are capability safe.
 
@@
 
+%\begin{theorem}\label{thm:world-existence}
+  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are preordered \cofes{}, and there exists an isomorphism $\xi$ such that
+  {\small 
+  \begin{align*}
+    \xi : \Wor & \cong \blater (\nats \finparfun \Regions)\\
+    \Regions & = \{\revoked\} \uplus \\
+             &\hskip -2cm \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
+            & \hskip -2cm \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
+ \end{align*}
+ }
+ and for $W, W' \in \Wor$
+ \begin{align*}
+   W \futurestr W' & \Leftrightarrow \xi(W) \futurestr \xi(W')   \\
+   W \futurewk W' & \Leftrightarrow \xi(W) \futurewk \xi(W')
+ \end{align*}
+%Let $\iota.v$ be the projection of the view of a region.
+Worlds are finite partial maps from region names to regions,
+and region names are simply natural numbers
+\begin{align*}
+  \Worlds & = \RegionNames \finparfun \Regions\\
+ \RegionNames & = \nats.
+\end{align*}
+The \emph{private future} region relation satisfies the following properties:
+\begin{mathpar}
+  \inferrule{  (s,s') \in \phi \\
+    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ r \in \Regions }
+  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ r \in \Regions }
+  { r \futurestr \revoked }
+\end{mathpar}
+The \emph{public future} region relations satisfies the following properties:
+\begin{mathpar}
+  \inferrule{  (s,s') \in \phi_\pub \\
+    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
+  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
+  \and
+  \inferrule{ }
+  { \revoked \futurewk \revoked }
+\end{mathpar}
+
+The two future world relations satisfy the following properties: They allow for any extension of the current world and all existing worlds are allowed to move to an appropriate future region. That is
+\begin{mathpar}
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
+  { W' \futurewk W }
+  \and
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
+  { W' \futurestr W }
+\end{mathpar}
+
 \section{Examples}

### 2017-02-05 csf-2017/paper.tex

@@
 \label{sec:introduction}
+%%% a)
+% Current low-level protection coarse grained memory protection or
+% properties of high-level languages are not really enforced?
+
+% Capability machines offer fine-grained memory protection
+A capability machine is a low-level machine that offers fine-grained
+memory protection.
+% Capability machines and unforgeable tokens for memory access. 
+This is done by introducing unforgeable tokens to
+the machine. These tokens are called capabillities and grant a kind of
+authority and a range of authority. We will refer to the capabilities
+that grant a combination of read, write, or execute permission as memory
+capabilities. Capabilities with the necessary permissions must be
+presented every time an instruction that manipulates the memory is
+executed.
+
+Memory capabilities provide memory protection, but they do not provide
+a way to set up security domains. When you jump to an untrusted piece
+of code that is supposed to jump back to you, then you have to either
+revoke all your capabilities or pass to the piece of code you do not
+trust. This short coming can be overcome by adding something like the
+enter capability from the M-Machine\todo{add reference}. The enter
+capability is an opaque capability which can only be used for a
+jump. When jumped to it grants read and execute permission which
+allows the code that is now executing to read capabilities stored in
+memory. 
+The CHERI processor's ccall achieves a similar property\todo{Add reference}.
+
+Capabilities are irrevocable, so when we pass a capability has to
+an untrusted program, then we have to assume that the program keeps it
+around indefenitely. This means that we cannot reuse the piece of
+memory that the capability governs essentially creating a memory
+leak. The CHERI processor has a special kind of capabilities called
+\emph{local capabilities}\todo{reference?} that introduces a simple
+kind of temporal information control which allows for a simple kind of
+capability revocation. This is achieved by adding a tag to every
+capability that marks it as either local or global. Local capabilities
+can only be written through a capability with a new permission called
+\emph{permit write local}.
+% More about local capabilities.
+
+With just memory capabilities, enter capabailities, and local
+capabilities, the capability machine is very powerful and can
+enforce properties that high-level languages promise. In most
+high-level languages, we expect local state to be private and
+encapsulated. The memory capabilities make sure that local state
+cannot be accessed without a capability to do so. With the enter
+capabaility, we can regain control of local state after passing
+control to an untrusted piece of code which gives us encapsulation
+when control is passed to an untrusted program.
+
+Many high-level languages guarantee that function function calls are
+well-bracketed. This can easily be achieved by having a trusted call
+stack. On a capability machine, it is possible to enforce well
+barcketedness without a trusted stack. When ensuring well-bracketed
+the two main challenges calls are 1) preventing storage of return
+pointers and 2) encapsulation of the local stack frame. If we don't
+have the first property, then an adversary can store the return
+capability in a call and jump to it in a later call. If the local
+stack frame is not properly encapsulated, then an adversary can break
+the well-bracketedness by jumping to a different calls return
+capability.
+
+In order to deal with the two challenges, we make sure that all return
+capabilities are local and that there are no global capabilities with
+permit write local permission. This makes sure that there is no way to
+save a return capability in persistent local state.\lau{this does not really make
+sense as the local stack frame can be seen as local state} It is,
+however, to limiting to not provide any means to store the return
+capabilities as a piece of code may need to store a return capability
+before jumping to an adversary. It is therefore needed to provide a
+local capability with permit write local permission which can be used
+as a stack. The stack should now be the only place one can store the
+return capability, but it should also be the only place one can store
+the stack capability.
+
+Using local return capabilities and putting a stack abstraction on a
+local capability permit write local capability is a good start, but as
+it turns out, it is not wnough. An adversary could do the following 
+\begin{enumerate}
+\item In a first invocation, fill the stack frame with copies of the
+  current stack capability.
+\item In a later call, the adversary may be able to load the stack
+  capability from the previous call.
+\item If we have used a larger part of the stack than in the first
+  call, then the adversary has access to part of our local stack frame.
+\end{enumerate}
+
+The above attack shows that it is tricky to ensure
+well-bracketedness. It raises the question whether it is possible to
+do so it is completely watertight. The attack utilized the stack, so
+it is necessary to have a trusted stack or can we do without? In
+general it raises the question, how do we reason about capability
+machines and local capabilities, so we are sure that what ever calling
+scheme we come up with actually provably ensures well-bracketed calls.
+
+In the following paper, we present the following contributions
+\begin{enumerate}
+\item 
+\end{enumerate}
+
+%%% b)
+
+% Examples of capability machines?
+
+% Enter capabilities
+
+% Local capabilities
+
+% High-level languages often provide guarantees such as encapsulation of private state and well-bracketedness of calls - often not clear how well this is enforced.
+Just enforced when interacting with other programs written in the same high-level language or is it also guaranteed when interacting with assembly programs.
+
+% High-level programs compiled to assembly can ensure these properties.
+
+%
 

### 2017-02-05 local-capability-machine.tex

--- 2017-02-03 local-capability-machine.tex
@@
   The theorem follows from a more general solution theorem for the
-  category of $P$ preordered \cofes, see~\cite{bizjak-note}. 
+  category of $P$ preordered \cofes,
+  see~\cite{Birkedal:2010:TCS:411:4102-4122}, \cite{Birkedal:tutorial-notes} and \cite{bizjak-note}. 
   We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to

### 2017-02-05 local-capability-machine.tex
Fixed the writeCond problem using a new notion of "address-stratified" regions.

--- 2017-02-04 local-capability-machine.tex
@@
       \item $\npair{(\start_\stk'', \addrend_\stk)} \in \writeCond{}(\iota^\pwl,\local)(W)$\\
-        This amounts to showing 
+        Using Lemma~\ref{lem:iota-pwl-address-stratified}, this amounts to showing
         \[
@@
         \item $\npair[n']{\start_\stk'', \addrend_\stk} \in \writeCond{}(\local)(W')$ \\
-          this amounts to showing
+          Using Lemma~\ref{lem:iota-pwl-address-stratified}, this amounts to showing
           \[
@@
     \item $\npair{(\start_\adv,\addrend_\adv)} \in \writeCond{}(\iota^\nwl,\glob)(W)$\\
-      Simple using region $3$ in $W$.
+      Simple using region $3$ in $W$, using
+      Lemma~\ref{lem:iota-nwl-address-stratified}.
   \end{enumproof}
@@
 We have omitted the details of this construction here, but it is done by parameterising all the definitions by the value relation.
+
+\begin{align*}
+&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
+&\qquad\begin{multlined}
+  \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
+  \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
+  \dom(\ms) = \dom(\ms') \wedge \\
+  \forall \addr \in
+  \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
+\end{multlined}
+\end{align*}
+
 \begin{align*}
@@
     & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \}
+    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{ and }\\
+    & \;\qquad W(r) \text{ is address-stratified } \}
   \end{aligned}
@@
 \begin{lemma}
+  \label{lem:iota-pwl-address-stratified}
+  $\iota^\pwl_{\start,\addrend}$ is address-stratified.
+\end{lemma}
+\begin{proof}
+  Easy unfolding of definitions.
+\end{proof}
+
+\begin{lemma}
   \label{lem:hnwl-mono-str}
@@
 \begin{lemma}
+  \label{lem:iota-nwl-address-stratified}
+  $\iota^\nwl_{\start,\addrend}$ is address-stratified.
+\end{lemma}
+\begin{proof}
+  Easy unfolding of definitions.
+\end{proof}
+
+\begin{lemma}
   \label{lem:iotanwlp-is-a-region}
@@
   \]
+  and 
+  \begin{equation*}
+    \revokeTemp(W)(r) \text{ is address-stratified}
+  \end{equation*}
   Notice $\revokeTemp{W}(r)$ is a $\perma$ region, so $\revokeTemp{W}(r) = W(r)$. Using $r$ as witness, the result is immediate.
@@
      [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
-     \iota^\pwl_{\start',\addrend'}$: by definition of $\writeCond{}$.
+     \iota^\pwl_{\start',\addrend'}$ and $W(r)$ is address-stratified: by definition of $\writeCond{}$.
    \item Suffices: $W(r)\nsupsim[n-1] \iota^\nwl_{\start',\addrend'}$. By
@@
  \begin{proof}
-   TODO
+   \begin{enumproof}
+   \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota,\gl)(W)$ where $\iota
+     = \iota^\pwl$ or $\iota = \iota^\nwl$ and (if $\var{w} =
+     ((\_,\local),\_,\_,\_)$, then $\iota = \iota^\pwl$).
+
+     By definition of $\stdvr(W)$ and $\writeAllowed{}$, from
+     $\npair{c}\in\stdvr(W)$, $((\perm,\gl),\start,\addrend,\addr) = c$ and
+     $\writeAllowed{\perm}$ and the fact that (if $\var{w} =
+     ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$)
+   \item $\exists r \in \var{localityReg}(\gl,W) \ldotp$ $\exists
+     [\start',\addrend'] \supseteq [\start,\addrend] \ldotp$ $W(r)\nsupsim[n-1]
+     \iota_{\start',\addrend'}$ and $W(r)$ is address-stratified. By definition
+     of $\writeCond{}$.
+   \item $\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp$
+     $\memSatPar{\ms'}{W}{P}$. By definition of $\heapSat[\ms']{n}{W}$.
+   \item $\ms' = \biguplus_{r\in\activeReg{W}}P(r)$ and $\forall r \in
+     \activeReg{W} \ldotp$ $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$ and
+     $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$. By definition of
+     $\memSatPar{\ms'}{W}{P}$.
+   \item $\exists H,s \ldotp$ $W(r) = (\_,s,\_,\_,H)$ and
+     $\npair[n]{P(r)} \in H(s)(\xi^{-1}(W))$. By instantiating the previous
+     point to the $r$ from the $\writeCond{}$.
+   \item $\npair{w} \in \iota.H~(\iota.s)~(\xi^{-1}(W))$ by definition of
+     $\iota^\pwl$, $\iota^\nwl$ and the fact that (if $\var{w} =
+     ((\_,\local),\_,\_,\_)$, then $\iota = \iota^\pwl$).
+   \item Define $\ms'_w$ such that $\dom(\ms'_w) = [\start',\addrend']$,
+     $\ms'_w(\addr) = w$ and $\ms'_w(\addr') = 0$ for $\addr' \neq \addr$. It's
+     easy to show from the previous point that $\npair{\ms'_w} \in
+     H(s)(\xi^{-1}(W))$.
+   \item $\dom(P(r)) = \dom(\ms'_w) = [\start',\addrend'] \ni \addr$ and
+     $\npair{P(r)\update{\addr}{w}} \in H(s)(\xi^{-1}(W))$ by applying the fact
+     that $W(r)$ is address-stratified, combined with the previous point.
+   \item Define $P'(r) = P(r)\update{a}{w}$ and $P'(r') = P(r')$ for $r' \neq r$.
+   \item $\ms'\update{a}{w} = \biguplus_{r\in\activeReg{W}}
+     P'(r)$ and $\ms'\update{a}{w} :_{n,P'} W$. By definition of
+     $\memSatPar{\ms'}{W}{P}$ and the previous two points.
+   \end{enumproof}
  \end{proof}
@@
     \item $\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\local)(W)$ : \label{pf:stack-cap-write}\\
-      Amounts to
+      Using Lemma~\ref{lem:iota-pwl-address-stratified}, this amounts to
       \[

### 2017-02-05 local-capability-machine.tex

@@
   category of $P$ preordered \cofes,
-  see~\cite{Birkedal:2010:TCS:411:4102-4122} and \cite{bizjak-note}. 
+  see~\cite{Birkedal:2010:TCS:411:4102-4122}, \cite{Birkedal:tutorial-notes} and \cite{bizjak-note}. 
   We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to

### 2017-02-06 csf-2017/paper.tex
added first high-level motivation

@@
 \label{sec:introduction}
+\todo[inline]{See the latex comments for the next paragraphs!}
+% Possible high-level intros:
+% 1) Security guarantees vs obstacles / arms race
+Security on processors is today subject to an arms race between
+processor manufacturers and hackers. On the hacker side, a constant
+effort is put into exploiting current processor designs and
+circumventing the security measures set in place to prevent known
+exploits. On the processor manufacturer side, new security measures
+are put in place to prevent new exploits. The arms race would come to
+a grinding halt if the security meassures provided more than obstacles
+and actually provided low-level security guarantees. One kind of
+low-level machines that provide low-level security is a capability machine.
+
+% 2) Memory access-control
+
+% 3) Fully abstract compilation
+
 %%% a)

### 2017-02-06 csf-2017/paper.tex
second high level motivation, secure compilation

@@
 % 3) Fully abstract compilation
+High-level languages provide guarantees such as encapsulation of local
+state and well-bracketedness of function calls. These guarantees are
+often relied upon when reasoning about the correctness and security of
+a program. High-level programs are compiled to machine code, so in
+order to rely on the correctness and security guarantees provided by
+the high-level language, it needs to be proven that the translation
+preserves these properties. Translated programs often interact with
+machine code that is not translated from the same high-level language,
+so in order to preserve the guarantees from the high-level languages,
+the low-level machine needs to provide some kind of security
+guarantees. In other words, a target of secure compilation needs to
+provide some security guarantees. Modern processors provide many
+security meassures, but very little in terms of guarantees. A machine
+that provides some means of security guarantees is a capability machine.
 

### 2017-02-06 csf-2017/paper.tex
changed ordering, added a comment
small intro update

@@
 % Possible high-level intros:
+% 0) Arms race compilers
+%
 % 1) Security guarantees vs obstacles / arms race
+\begin{comment}
   Security on processors is today subject to an arms race between
@@
   exploits. On the processor manufacturer side, new security measures
-are put in place to prevent new exploits. The arms race would come to
-a grinding halt if the security meassures provided more than obstacles
-and actually provided low-level security guarantees. One kind of
-low-level machines that provide low-level security is a capability machine.
-
-% 2) Memory access-control
-
-% 3) Fully abstract compilation
+  are put in place to prevent new exploits. The arms race would come
+  to a grinding halt if the security meassures provided more than
+  obstacles and actually provided low-level security guarantees. One
+  kind of low-level machines that provide low-level security is a
+  capability machine.
+\end{comment}
+% 2) Fully abstract compilation
 High-level languages provide guarantees such as encapsulation of local
@@
 
+% 3) Memory access-control
+% I don't like this angle and can't find a good way to sell it.
+
+% 4) History
+% Proposals throughout history.
+% Lack of formal account of properties-guarantees.
+
 %%% a)
@@
 % Capability machines offer fine-grained memory protection
+% Too technical
 A capability machine is a low-level machine that offers fine-grained
@@
 the machine. These tokens are called capabillities and grant a kind of
-authority and a range of authority. We will refer to the capabilities
+authority and a range of authority. 
+% Dynamic checks encapsulation, formal model
+
+ We will refer to the capabilities
 that grant a combination of read, write, or execute permission as memory
-capabilities. Capabilities with the necessary permissions must be
+capabilities. 
+ Capabilities with the necessary permissions must be
 presented every time an instruction that manipulates the memory is
@@
 leak. The CHERI processor has a special kind of capabilities called
-\emph{local capabilities}\todo{reference?} that introduces a simple
+\emph{local capabilities}\todo{reference?}\ that introduces a simple
 kind of temporal information control which allows for a simple kind of

### 2017-02-06 local-capability-machine.tex

@@
               \begin{specification}[Malloc v.2]
+                \label{spec:malloc}
                 $c_\malloc$ satisfies the specification for malloc iff
@@
                   \begin{aligned}
-                    &c_\malloc = ((\entry,\_),\_,\_,\_) \land\\
+                    &c_\malloc = ((\entry,\glob),\_,\_,\_) \land\\
                     &\exists \iota_{\malloc,0} \ldotp \\
-                    &\quad (\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (W) = \iota'.H (\iota'.s) ([i \mapsto W(i)]) ) \; \land \\
+                    &\quad (\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)])) ) \; \land \\
                     &\quad \iota_{\malloc,0}.v = \perma \; \land \\
                     &\quad (\forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
-                    &\qquad \forall i, n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \\
+                    &\qquad \forall i, n, \size \in \nats \ldotp \forall
+                    w_{\var{ret}} \in \Words \ldotp \\
                     &\qquad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
-                    &\qquad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land\\
-                    &\qquad \quad \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
-                    &\qquad \quad \memreg(r_1) = \size \land \\
-                    &\qquad \quad \size \geq 0 \land \\
-                    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
+                    &\qquad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
+                    &\qquad \quad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
                     &\qquad \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
@@
                     &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-                    &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-                    &\qquad \qquad \qquad \size - 1 = e'-b' )
+                    &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+                    &\qquad \qquad \qquad \size - 1 = e'-b' ) \land\\
+                    &\quad (\forall \Phi \in \ExecConfs \ldotp (\memreg(r_1)
+                    \not\in \ints \vee \memreg(r_1) < 0) \land \memreg(\pcreg) = \updatePcPerm{c_\malloc} \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed)
                   \end{aligned}
@@
   \label{lem:malloc-in-vr}
-  If $W(r) = \iota$ with $\iota \futurestr \iota_{\malloc,0}$, then
+  If $W(r) \futurestr \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
@@
 \begin{proof}
-  TODO
+  \begin{enumproof}
+  \item $c_\malloc = ((\entry,\glob),\base,\addrend,\addr)$.\\
+    By: the malloc specification (Specification~\ref{spec:malloc}).
+  \item Suffices: $\npair{(\start,\addrend,\addr)}\in\entryCond{\glob}(W)$.\\
+    By definition of $\stdvr(W)$.
+  \item Assume: $n' < n$, $W'\futurestr W$.\\
+    Suffices: $\npair[n']{((\exec,\glob),\start,\addrend,\addr)} \in
+    \stder(W')$.\\
+    By: definition of the $\entryCond{}$
+  \item Assume: $n''\leq n'$,
+    $\npair[n'']{\reg}\in\stdrr(W')$, $\heapSat[\ms]{n''}{W'}$\\
+    Suffices:
+    $\npair[n'']{(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms)}\in\observations(W')$\\
+    By: definition of $\stder(W')$
+  \item Assume: $i < n''$,
+    $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)\step[i]
+    (\halted,\mem')$\\
+    Suffices: $\exists W''\futurestr W',\ms_r,\ms'\ldotp$ $\mem' =
+    \ms'\uplus\ms_r\uplus\ms_f$ and $\heapSat[\ms']{n-i}{W''}$\\
+    By: definition of $\observations(W')$
+  \item $W'(r) \futurestr \iota_{\malloc,0}$\\
+    Easy from: $W'\futurestr W$ and $W(r) \futurestr \iota_{\malloc,0}$ using
+    transitivity of $\futurestr$.
+  \item $\exists P : \activeReg{W'} \rightarrow \HeapSegments \ldotp$
+    $\memSatPar[n'']{\ms}{W'}{P}$, i.e. $\hs = \biguplus_{r\in\activeReg{W'}}P(r)$ and
+    $\forall r \in \activeReg{W'} \ldotp$ $\exists H,s \ldotp$ $W'(r) =
+    (\_,s,\_,\_,H)$ and $\npair[n'']{P(r)} \in H(s)(\xi^{-1}(W'))$\\
+    By: definition of $\heapSat[\ms]{n''}{W'}$.
+  \item Define $\ms_{\var{frame}} = \left(\biguplus_{r'\in\activeReg{W'},r'\neq
+        r}P(r')\right)\uplus\ms_f$. Then $\ms\uplus\ms_f = P(r) \uplus
+    \ms_{\var{frame}}$ and $\heapSat[P(r)]{n''}{W'(r).H~(W'(r).s)~(\xi^{-1}(W'))}$.
+    Easy from the previous point.
+  \item $\heapSat[P(r)]{n''}{W'(r).H~(W'(r).s)~(\xi^{-1}([r\mapsto W'(r)]))}$, i.e.
+    $\heapSat[P(r)]{n''}{[r\mapsto W'(r)]}$.\\
+    By: the malloc specification (Specification~\ref{spec:malloc}) from the
+    previous point.
+  \item Case: $\reg(r_1) \in \ints$ and $\reg(r_1) \geq 0$
+    \begin{enumproof}
+    \item Define $\var{size} = \reg(r_1)$
+    \item $\exists \Phi' \in \ExecConfs, \ms_{\var{footprint}}',
+      \ms_{\var{alloc}} \in \HeapSegments, j \in \nats, b',e'\in \Addrs,
+      \iota_\malloc' \in \Regions \ldotp$
+      $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)
+      \step[j] \Phi'$ and $\memheap[\Phi']=\ms_{\var{footprint}}' \uplus
+      \hs_{\var{alloc}} \uplus \ms_{\var{frame}}$ and $\iota_{\malloc}'
+      \futurewk W'(r)$ and $\heapSat[\ms_{\var{footprint}}']{n-j}{[r \mapsto
+        \iota_\malloc']}$ and $\dom(\hs_{\var{alloc}}) = [b',e']$ and $\forall a
+      \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0$ and $\memreg[\Phi'] =
+      \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')}$
+      and $\var{size} - 1 = e'-b' )$.\\
+      By: the malloc specification (Specification~\ref{spec:malloc}).
+    \item Define $W'' = W'\update{r}{\iota_\malloc'}[i\mapsto
+      \iota^\nwl_{b',e'}]$ for $i \not\in\dom(W')$
+    \item \dominique{TODO}
+    \end{enumproof}
+  \item Case: $\reg(r_1)\not\in\ints \vee \reg(r_1) < 0$
+    \begin{enumproof}
+    \item $\exists j\ldotp (\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)\step[j]
+      \failed$\\
+      By: the malloc specification (Specification~\ref{spec:malloc}).
+    \item Contradiction with 
+      $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)\step[i]
+      (\halted,\mem')$
+    \end{enumproof}
+  \end{enumproof}
 \end{proof}

### 2017-02-06 csf-2017/paper.tex
notation in world theorem
intuition about logical relation

@@
 
+\emph{adversary}: It doesn't sound ``sexy'' to constantly say ``an untrusted
+piece of code'' should we adopt the convention that adversary refers
+to an unstrusted piece of code?
+
 \section{Introduction}
@@
 \section{Logical relation}
+In order to be able to reason about programs running on the capability
+machine and prove properties such as well-bracketedness and
+encapsulation about them, we define a unary logical relation.
+%
+% Intuition
+The logical relation is a semantic models of all the words on the
+capability machine that are definitely safe to give to an
+adversary. Whether a capability is safe to give to an adversary
+depends heavily on the contents of the memory as instructions as well
+as capabilities may reside in memory. The memory will be modeled using
+worlds where a world can be seen as a collection of protocols that
+parts of the memory must satisfy. The memory changes during an
+execution as memory is allocated and part of memory changes
+purpose. The model needs to be able to express these changes to the
+memory which is captured by the future world relations.
+%
 
-\begin{itemize}
-\item Intuition
-\item Worlds
-\item Logical relation
-\item Fundamental theorem
-\item Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer
-\end{itemize}
-
+\subsection{Worlds}
 %\begin{theorem}\label{thm:world-existence}
@@
  \begin{align*}
-   W \futurestr W' & \Leftrightarrow \xi(W) \futurestr \xi(W')   \\
-   W \futurewk W' & \Leftrightarrow \xi(W) \futurewk \xi(W')
+   W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
+   W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
@@
 \end{mathpar}
+
+\subsection{Logical relation}
+
+\subsection{Fundamental theorem of logical relations}
+
+\subsubection{Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
 
@@
 
+\bibliographystyle{plainnat}
+\bibliography{references}
 \end{document}

### 2017-02-06 csf-2017/paper.tex
Added some subsections to logical relation section

@@
 
-\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
-\declaretheorem[numbered=yes,name=Theorem,qed=$\blacksquare$]{theorem}
-\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
-\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}
+\newtheorem{theorem}{Theorem}
 
@@
 \subsection{Worlds}
-%\begin{theorem}\label{thm:world-existence}
-  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are preordered \cofes{}, and there exists an isomorphism $\xi$ such that
+For this logical relation, we use step-indexed Kripke worlds. They are
+defined by the $\Regions$ given in
+
+\begin{theorem}\label{thm:world-existence}
+  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and
+  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
+  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
   {\small
@@
     \end{align*}
- }
- and for $W, W' \in \Wor$
+  } and for $W, W' \in \Wor$
   \begin{align*}
@@
   \end{align*}
+\end{theorem}
+
 %Let $\iota.v$ be the projection of the view of a region.
@@
 \end{align*}
+
+\subsubsection{Future worlds}
+
+{\bf Public future worlds}
+The \emph{public future} region relations satisfies the following properties:
+\begin{mathpar}
+  \inferrule{  (s,s') \in \phi_\pub \\
+    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
+  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
+  \and
+  \inferrule{ }
+  { \revoked \futurewk \revoked }
+\end{mathpar}
+
+{\bf Private future worlds}
+
 The \emph{private future} region relation satisfies the following properties:
@@
   { r \futurestr \revoked }
-\end{mathpar}
-The \emph{public future} region relations satisfies the following properties:
-\begin{mathpar}
-  \inferrule{  (s,s') \in \phi_\pub \\
-    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
-  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
-  \and
-  \inferrule{ }
-  { \revoked \futurewk \revoked }
 \end{mathpar}
@@
 
-\subsubection{Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
+\subsection{Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
 

### 2017-02-06 local-capability-machine.tex
notation in world theorem
Prove that malloc is safe to pass to an adversary.

@@
 
+\begin{lemma}[Determinacy]
+  \label{lem:determinacy}
+  If $\Phi \step \Phi'$ and $\Phi \step \Phi''$, then $\Phi' = \Phi''$.
+  If $\Phi \step[n] \Phi'$ and $\Phi \step[n] \Phi''$, then $\Phi' = \Phi''$.
+  If $\Phi \step[n] \Phi'$ and $\Phi \step[n'] (\halted,\mem'')$, then $n \geq n'$
+  and $\Phi' \step[n'-n] (\halted,\mem'')$.
+\end{lemma}
+\begin{proof}
+  By easy inspection of the definition of the operational semantics.
+\end{proof}
+
 \section{Malloc specification}
@@
                     &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-                    &\qquad \qquad \quad \exists j \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+                    &\qquad \qquad \quad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
                     &\qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
@@
 
-        \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: for $n'' <
-          n'-1$ and $W' \futurestr W$ show:
-          \\$\npair[n'']{\updatePcPerm{c_\adv}} \in \stder(W')$.
-          Use FTLR. \label{f1:adv}
+        \item $\npair[n'-1]{c_\adv} \in \stdvr(W)$: for $n'' < n'-1$ and $W'
+          \futurestr W$ show: \\$\npair[n'']{\updatePcPerm{c_\adv}} \in
+          \stder(W')$. Follows from Lemma~\ref{lem:safe-values-safe-invoke},
+          together with Lemma~\ref{lem:stdvr-glob-priv-mono} and the fact that
+          $c_\adv$ is non-local.  \label{f1:adv}
         \end{enumproof}
@@
 \begin{align*}
-  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+  & \writeCond{} : (((\Addrs\times\Addrs) \fun\Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
   & \writeCond{}(\iota,\gl)(W) =  \\
@@
 
+\begin{lemma}[Safe values are safe to invoke.]
+  \label{lem:safe-values-safe-invoke}
+  If $\npair[n+1]{w} \in \stdvr(W)$, then $\npair{\updatePcPerm{w}} \in \stder(W)$.
+\end{lemma}
+\begin{proof}
+  \begin{enumproof}
+  \item Case $w = \stdcap$ and $\start \leq \addr \leq \addrend$
+    and $\perm \in \{ \exec,\rwx, \rwlx \}$:
+    \begin{enumproof}
+    \item $\npair[n+1]{(\perm,\start,\addrend)}\in\execCond{\gl}(W)$.\\
+      By: definition of $\stdvr(W)$ using the fact that $\perm \in \{ \exec,\rwx, \rwlx \}$.
+    \item $\npair[n]{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$:
+      By definition of $\execCond{}$ using the fact that $\start \leq \addr \leq \addrend$.
+    \end{enumproof}
+  \item Case $w = \stdcap$ and $\start \leq \addr \leq \addrend$
+    and $\perm = \entry$:
+    \begin{enumproof}
+    \item $\npair[n+1]{(\start,\addrend,\addr)}\in\entryCond{\gl}(W)$.\\
+      By: definition of $\stdvr(W)$ using the fact that $\perm = \entry$.
+    \item $\npair[n]{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W)$:
+      By definition of $\entryCond{}$ using the fact that $\start \leq \addr \leq \addrend$.
+    \item $\updatePcPerm{w} = ((\exec,\gl),\start,\addrend,\addr)$:\\
+      By definition of $\updatePcPerm{\cdot}$
+    \end{enumproof}
+  \item Otherwise:
+    $\npair{\updatePcPerm{w}} \in \stder(W)$:\\
+    By Lemma~\ref{lem:failed-obs-stder}.
+  \end{enumproof}
+\end{proof}
+
 \begin{lemma}[Malloc is safe to pass to adversary]
@@
     \item $\exists \Phi' \in \ExecConfs, \ms_{\var{footprint}}',
-      \ms_{\var{alloc}} \in \HeapSegments, j \in \nats, b',e'\in \Addrs,
-      \iota_\malloc' \in \Regions \ldotp$
+      \ms_{\var{alloc}} \in \HeapSegments, j \in \nats, j > 0 \land b',e'\in
+      \Addrs, \iota_\malloc' \in \Regions \ldotp$
       $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)
@@
       \hs_{\var{alloc}} \uplus \ms_{\var{frame}}$ and $\iota_{\malloc}'
-      \futurewk W'(r)$ and $\heapSat[\ms_{\var{footprint}}']{n-j}{[r \mapsto
+      \futurewk W'(r)$ and $\heapSat[\ms_{\var{footprint}}']{n''-j}{[r \mapsto
         \iota_\malloc']}$ and $\dom(\hs_{\var{alloc}}) = [b',e']$ and $\forall a
@@
       \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')}$
-      and $\var{size} - 1 = e'-b' )$.\\
+      and $\var{size} - 1 = e'-b' )$ with $w_{\var{ret}} = \memreg(r_1)$.\\
     By: the malloc specification (Specification~\ref{spec:malloc}).
     \item Define $W'' = W'\update{r}{\iota_\malloc'}[i\mapsto
-      \iota^\nwl_{b',e'}]$ for $i \not\in\dom(W')$
-    \item \dominique{TODO}
+      \iota^\nwl_{b',e'}]$ for $i \not\in\dom(W')$. We have that $W''\futurewk
+      [r\mapsto\iota_\malloc']$ and $W'' \futurewk W'$.\\
+      By: definition of $\futurewk$ , using the fact that
+      $\iota_\malloc'\futurewk W(r)$.
+    \item $\npair[n''']{(\start',\addrend')}\in\readCond{\glob}(W'')$ for all $n'''$:\\
+      By: definition of $\readCond{}$, using the region $W''(i)$ and Lemma~\ref{lem:nwl-subset-pwl}.
+    \item $\npair[n''']{(\start',\addrend')}\in\writeCond{\iota^\nwl,\glob}(W'')$ for all $n'''$:\\
+      By: definition of $\writeCond{}$, using the region $W''(i)$.
+    \item
+      $\npair[n''']{(p,\start',\addrend')}\in\execCond{\iota^\nwl,\glob}(W'')$
+      for all $n'''$, $p \in \{\rwx,\exec\}$:\\
+      By: the definition of $\execCond{}, $ the FTLR (Theorem~\ref{thm:ftlr})
+      using Lemmas~\ref{lem:writecond-mono-priv}, \ref{lem:readcond-mono-priv}
+      and the previous two points.
+    \item $\npair[n'']{((\rwx,\glob),b',e',b')}\in\stdvr(W'')$:\\
+      By: definition of $\stdvr(W'')$ and the above three points.
+    \item
+      $\npair[n''-j]{\memreg[\Phi]\update{r_1}{((\rwx,\glob),b',e',b')}}\in\stdrr(W'')$:\\
+      By Lemma~\ref{lem:stdrr-dc}, Lemma~\ref{lem:stdrr-mono} using the fact
+      that $W''\futurewk W'$ and $\npair[n'']{\memreg}\in\stdvr(W')$, together
+      with the previous point.
+    \item $\npair[n''']{\ms_{\var{alloc}}} \in \iota^\nwl_{b',e'}.H~
+      \iota^\nwl_{b',e'}.s~ W''$ for any $n'''$:\\
+      By definition of $\iota^\nwl$, $H^\nwl$ and $\stdvr(\cdot)$ and the facts
+      that $\dom(\hs_{\var{alloc}}) = [b',e']$ and $\forall a
+      \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0$.
+    \item Define $\ms' = \left(\biguplus_{r'\in\activeReg{W'},r'\neq
+          r}P(r')\right)\uplus \ms_{\var{footprint}}'\uplus\ms_{\var{alloc}}$.
+      Then
+      $\memheap[\Phi'] = \ms' \uplus \ms_f$ and $\heapSat[\ms']{n''-j}{W''}$:\\
+      By the facts that $\memheap[\Phi']=\ms_{\var{footprint}}' \uplus
+      \hs_{\var{alloc}} \uplus \ms_{\var{frame}}$, $\ms_{\var{frame}} =
+      \left(\biguplus_{r'\in\activeReg{W'},r'\neq r}P(r')\right)\uplus\ms_f$,
+      the previous point, the facts that
+      $\heapSat[\ms_{\var{footprint}}']{n''-j}{[r \mapsto \iota_\malloc']}$ and
+      $W''\futurewk [r\mapsto\iota_\malloc']$, the facts that ($\forall r \in
+      \activeReg{W'} \ldotp$ $\exists H,s \ldotp$ $W'(r) =, (\_,s,\_,\_,H)$ and
+      $\npair[n'']{P(r)} \in H(s)(\xi^{-1}(W'))$) and $W'' \futurewk W'$ and the
+      public monotonicity and downwards closedness of all islands, and finally
+      the definition of $W''$.
+    \item $\npair[n''-j+1]{w_{\var{ret}}} \in \stdvr(W'')$:\\
+      By Lemma~\ref{lem:stdvr-mono-wk}, the fact that $W''\futurewk W'$,
+      Lemma~\ref{lem:stdvr-dc}, and teh fact that
+      $\npair[n'']{w_{\var{ret}}}\in\stdvr(W')$, which follows from
+      $w_{\var{ret}} = \memreg(r_1)$ and $\npair[n'']{\reg}\in\stdrr(W')$.
+    \item $\npair[n''-j]{\updatePcPerm{w_{\var{ret}}}} \in \stder(W'')$:\\
+      By Lemma~\ref{lem:safe-values-safe-invoke} from the previous point.
+    \item
+      $\npair[n''-j]{(\memreg[\Phi]\update{r_1}{((\rwx,\glob),b',e',b')}\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}},
+        \ms')} \in \observations(W'')$:\\
+      By: definition of $\stder(W'')$, using the previous point and the facts
+      that
+      $\npair[n''-j]{\memreg[\Phi]\update{r_1}{((\rwx,\glob),b',e',b')}}\in\stdrr(W'')$,
+      $\heapSat[\ms']{n''-j}{W''}$
+    \item $i > j$ and $\Phi'\step[i-j]
+      (\halted,\mem')$.\\
+      By combining
+      $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)\step[i]
+      (\halted,\mem')$ with
+      $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)
+      \step[j] \Phi'$ using Lemma~\label{lem:determinacy}.
+    \item $\exists W''' \futurestr W'', \hs_r, \hs'' \ldotp$ $\heap' = \hs'' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs'']{n-i}{W'''}$.\\
+      By: definition of $\observations(W''')$ from the two previous points.
+    \item $W''' \futurestr W'$: \\
+      By Lemma~\ref{lem:future-priv-pub-trans}, using the previous point and the fact that $W'' \futurewk
+      W'$.
     \end{enumproof}
@@
 \begin{lemma}[Transitivity proberties between private and public future worlds]
+  \label{lem:future-priv-pub-trans}
 \[
@@
 \subsubsection{Value relation}
+\begin{lemma}[Value relation downwards closed]
+  \label{lem:stdvr-dc}
+  \[
+    n' \leq n \land \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
+  \]
+\end{lemma}
+\begin{proof}
+  By definition of $\stdvr(W)$ using Lemma~\ref{lem:readcond-dc},
+  \ref{lem:writecond-dc}, \ref{lem:execcond-dc} and \ref{lem:entrycond-dc}.
+\end{proof}
+
+\begin{lemma}[Register relation downwards closed]
+  \label{lem:stdrr-dc}
+  \[
+    n' \leq n \land \npair{w} \in \stdrr(W) \Rightarrow \npair[n']{w} \in \stdrr(W)
+  \]
+\end{lemma}
+\begin{proof}
+  By definition of $\stdrr(W)$ using Lemma~\ref{stdvr-dc}.
+\end{proof}
+
 \begin{lemma}[Value relation monotone wrt $\futurewk$]

### 2017-02-06 csf-2017/paper.tex

@@
 encapsulation about them, we define a unary logical relation.
-%
+
 % Intuition
-The logical relation is a semantic models of all the words on the
+The logical relation is a semantic model of all the words on the
 capability machine that are definitely safe to give to an
 adversary. Whether a capability is safe to give to an adversary
-depends heavily on the contents of the memory as instructions as well
+depends heavily on the contents of the memory since instructions as well
 as capabilities may reside in memory. The memory will be modeled using
-worlds where a world can be seen as a collection of protocols that
-parts of the memory must satisfy. The memory changes during an
-execution as memory is allocated and part of memory changes
-purpose. The model needs to be able to express these changes to the
-memory which is captured by the future world relations.
+worlds. A world consists of a finite set of regions, and each region
+can be understood as a protocol describing the possible evolution of
+part of the memory. Memory changes during an
+execution since new memory can be allocated and part of the existing
+memory can change. We use two future world relations to capture
+such changes to the memory in our model.
 %

### 2017-02-07 csf-2017/paper.tex
Tune math sizes in the same way as for the EuroS&P2016 paper.
region intuition

@@
 
-\subsection{Worlds}
-For this logical relation, we use step-indexed Kripke worlds. They are
-defined by the $\Regions$ given in
+\subsection{Regions}
+%Regions intuition
+We have three types of regions: $\perma$, $\temp$, and $\revoked$. The
+$\perma$ and $\temp$ regions contains a state transition system with
+public and private transitions as known from logical relations
+from high-level languages\todo{reference}. The state transition
+systems can be understood as protocols the memory must follow. The
+protocols imposed on memory by $\perma$ regions stay in place
+indefinitely. This means that any capability, $\local$ or $\glob$, can
+depend on these protocols. The protocols imposed on memory by $\temp$
+regions can be revoked at certain points during execution
+corresponding to the points when $\local$ capabilities can be
+revoked. This means that $\local$ capabilties safely can depend on the
+protocols imposed on memory by $\temp$ regions, but $\glob$
+capabilities cannot as the $\glob$ capability may stay around
+throughout the execution of a program, but the $\temp$ regions may not
+stay around.
+
+For monotonicity reasons, a region cannot disappear, but
+we need a way to remove $\temp$ regions. This is where the $\revoked$ region comes
+into the picture. The $\revoked$ region contains no state transition
+system and puts no requirements on the memory it models. It serves
+simply as a mask for the $\temp$ regions that have been revoked.
 
@@
 
+\subsection{Worlds}
+For this logical relation, we use step-indexed Kripke worlds. They are
+defined by the $\Regions$ given in
+
 %Let $\iota.v$ be the projection of the view of a region.
@@
 \subsubsection{Future worlds}
+There are two future world relations a \emph{public future world} relation and a \emph{private future world} relation.
+In both future world relations, we have extension ordering, so a future world $W'$ of $W$ has at least the same regions as $W$, but it may add new regions imposing new protocols on part of the memory.
+The two future world relations differ in the way existing regions are allowed to evolve.
+In a \emph{public future region}
+\begin{mathpar}
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
+  { W' \futurewk W }
+  \and
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
+  { W' \futurestr W }
+\end{mathpar}
 
-{\bf Public future worlds}
+{\bf Public future regions}
 The \emph{public future} region relations satisfies the following properties:
@@
 
-{\bf Private future worlds}
-
+{\bf Private future regions}
 The \emph{private future} region relation satisfies the following properties:
@@
 The two future world relations satisfy the following properties: They allow for any extension of the current world and all existing worlds are allowed to move to an appropriate future region. That is
-\begin{mathpar}
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
-  { W' \futurewk W }
-  \and
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
-  { W' \futurestr W }
-\end{mathpar}
 

### 2017-02-07 csf-2017/paper.tex
recursive world intuition and a bit technical stuff

@@
 
+% domi: IEEEtran.cls sets font size to 10bp, apparently.
+% not sure what the difference is with 10pt.
+\DeclareMathSizes{10bp}{8}{7}{6}
+
+%\def\IEEEbibitemsep{0pt}
+
+\makeatletter
+\setlength\mpr@andskip{1em}
+\def\mpr@lineskip{5em}
+\def\MathparLineskip{\mpr@lesslineskip}
+\makeatother
+
+\newlength{\oldmathindent}
+\newenvironment{withmathindent}[1]{\setlength{\oldmathindent}{\mathindent}\setlength{\mathindent}{#1}}{\setlength{\mathindent}{\oldmathindent}}
+
 \begin{document}
+\setlength{\mathindent}{.2cm}
 
@@
 
-\subsection{Regions}
-%Regions intuition
+\subsection{Worlds}
+% Intuition worlds
+The worlds are step-indexed Kripke worlds. For the reader not familiar
+with these, it is sufficient to think of worlds as finite sets of
+regions.
+
+% Regions intuition
 We have three types of regions: $\perma$, $\temp$, and $\revoked$. The
@@
 
+% Recursive worlds intuition
+The worlds end up being recursive. The memory can contain capabilities
+and capabilities grants authority over part of the memory. A world
+models memory, so it also needs to be able to model capabilities in
+memory and the protocols on them. Whether a capability satisfies the
+protocol of a region may depend on what it grants authority over in
+memory which is described by the world.
+
+% Short technical explanation
+The recursiveness of the worlds gives rise to a recurisve domain
+equation. Theorem~\ref{thm:world-existence} gives the solution to this
+recursive domain equation.
 \begin{theorem}\label{thm:world-existence}
@@
 \end{theorem}
-
-\subsection{Worlds}
-For this logical relation, we use step-indexed Kripke worlds. They are
-defined by the $\Regions$ given in
-
-%Let $\iota.v$ be the projection of the view of a region.
-Worlds are finite partial maps from region names to regions,
-and region names are simply natural numbers
+We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
+the worlds:
 \begin{align*}
   \Worlds & = \RegionNames \finparfun \Regions\\
- \RegionNames & = \nats.
+ \RegionNames & = \nats
 \end{align*}

### 2017-02-07 csf-2017/paper.tex
private future region relation

@@
 \subsection{Worlds}
+\todo[inline]{Find a good place to describe monotonicity of
+  interpretation function.}
 % Intuition worlds
@@
 with these, it is sufficient to think of worlds as finite sets of
-regions.
+regions that impose protocols on the memory.
 
@@
 \subsubsection{Future worlds}
-There are two future world relations a \emph{public future world} relation and a \emph{private future world} relation.
-In both future world relations, we have extension ordering, so a future world $W'$ of $W$ has at least the same regions as $W$, but it may add new regions imposing new protocols on part of the memory.
-The two future world relations differ in the way existing regions are allowed to evolve.
-In a \emph{public future region}
+In order to model changes to memory over time, we use the future world
+relations. The future world relations model how more memory may be
+allocated but also how memory may change. How memory is allowed to
+change over time depends on the region that governs it and the future
+world relation used. The most interesting part about the future worlds
+is how it affects the regions, so we first look at the properties of
+the \emph{public future region} relation and the \emph{private future
+  region} relation. We then use these relations to describe the future
+world relations.
+
+{\bf Private future regions}
+In the private future region relation, we allow revocation of $\temp$
+regions. The regions cannot be revoked by removing them from the
+world, because we need the future world relations to be
+monotone. This is why we have the $\revoked$ region, so we have
+a region that can serve as a mask. We can, however, let any region
+serve as a mask, so in the private future region relation, we allow a
+$\temp$ region to become any valid region. As the $\revoked$ region
+serves as a mask, we let any valid region take its place in the
+private future region relation.
+
+If a $\temp$ region is not masked by a region, then it behaved like a
+$\perm$ region. The $\perma$ regions stay the same type and they are
+allowed to take a transition in the private part of the transition
+system. Which can be thought of as the memory changes according to
+some protocol.
+
+The properties of the \emph{private future region} relation are:
 \begin{mathpar}
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
-  { W' \futurewk W }
+  \inferrule{  (s,s') \in \phi \\
+    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
   \and
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
-  { W' \futurestr W }
+  \inferrule{ r \in \Regions }
+  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ r \in \Regions }
+  { r \futurestr \revoked }
 \end{mathpar}
@@
 {\bf Public future regions}
+The public future regions relation has the following properties:
+% - permanent and temporary regions stay the same type
+% - public transitions
+The $\perma$ and $\temp$ regions stay the same type of region and they
+are allowed to take a transition in the public part of their
+transition systems.
+
+% - revoked regions can be reused as temporary regions
+The $\revoked$ regions can either stay $\revoked$, or they can become
+any $\temp$ region. When a region is $\revoked$ it serves as a mask
+for a revoked $\temp$ region. We use masks rather than deleting the
+region altogether for monotonicity reasons, but we might as well let a
+new $\temp$ region serve as the mask. We do not allow $\revoked$
+regions to become a $\perma$ region because we want the private future
+region relation to be able to reuse any region it has revoked.
+
 The \emph{public future} region relations satisfies the following properties:
@@
 
-{\bf Private future regions}
-The \emph{private future} region relation satisfies the following properties:
+% The future world relation are really the ones from thm, the above
+% are properties of them.
+{\bf Future world relations}
+There are two future world relations a \emph{public future world}
+relation and a \emph{private future world} relation. In both future
+world relations, we have extension ordering, so a future world $W'$ of
+$W$ has at least the same regions as $W$, but $W'$ may have new regions
+imposing new protocols on parts of the memory. The two future world
+relations differ in the way existing regions are allowed to evolve. In
+a \emph{public future world}, the regions are allowed to evolve
+according to the \emph{public future region} relation. In a
+\emph{private future world}, the regions are allowed to evolve
+according to the \emph{private future world} relation. The future
+world relations are defined as follows:
 \begin{mathpar}
-  \inferrule{  (s,s') \in \phi \\
-    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
+  { W' \futurewk W }
   \and
-  \inferrule{ r \in \Regions }
-  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ r \in \Regions }
-  { r \futurestr \revoked }
+  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
+  { W' \futurestr W }
 \end{mathpar}
 
-The two future world relations satisfy the following properties: They allow for any extension of the current world and all existing worlds are allowed to move to an appropriate future region. That is
+% Relate to STSs for high-level languages. Have the public-private
+% transitions, but they are also used for more than this
 

### 2017-02-07 csf-2017/paper.tex
public future region

@@
 {\bf Public future regions}
-The public future regions relation has the following properties:
-% - permanent and temporary regions stay the same type
-% - public transitions
-The $\perma$ and $\temp$ regions stay the same type of region and they
-are allowed to take a transition in the public part of their
-transition systems.
+In the public future regions we do not permit revocation of $\temp$
+regions. In this relation, the $\temp$ and $\perma$ regions must stay
+the same type and they are allowed to take a transition in the public
+part of the transition system.
 
-% - revoked regions can be reused as temporary regions
-The $\revoked$ regions can either stay $\revoked$, or they can become
-any $\temp$ region. When a region is $\revoked$ it serves as a mask
-for a revoked $\temp$ region. We use masks rather than deleting the
-region altogether for monotonicity reasons, but we might as well let a
-new $\temp$ region serve as the mask. We do not allow $\revoked$
-regions to become a $\perma$ region because we want the private future
-region relation to be able to reuse any region it has revoked.
+The $\revoked$ regions serve as masking regions, but are not as
+liberal as in the private future region relation when it comes to
+letting other regions replace them as masks. In the public future
+region relation, we only allow $\revoked$ and $\temp$ regions to take
+on the job as masking region. We have this limitation, because we
+want the private future region relation to be able to reuse any region
+it has revoked.
 
@@
 \end{mathpar}
+
+Notice that the public future region relation is a subset of the
+private future region relation, so anything you are allowed to do in
+the public future region relation, the private future region relation
+can do the same.
 

### 2017-02-07 csf-2017/paper.tex
public future regions

@@
 imposing new protocols on parts of the memory. The two future world
-relations differ in the way existing regions are allowed to evolve. In
-a \emph{public future world}, the regions are allowed to evolve
-according to the \emph{public future region} relation. In a
-\emph{private future world}, the regions are allowed to evolve
-according to the \emph{private future world} relation. The future
-world relations are defined as follows:
+relations differ in the way existing regions are allowed to
+evolve. The public future world relation uses the public future region
+relation and the private future world relation uses the private future
+region relation:
 \begin{mathpar}
@@
 % transitions, but they are also used for more than this
+\todo[inline]{Relate to STSs for high-level languages?}
 

### 2017-02-07 csf-2017/paper.tex
lr editing

@@
 depends heavily on the contents of the memory since instructions as well
-as capabilities may reside in memory. The memory will be modeled using
-worlds. A world consists of a finite set of regions, and each region
+as capabilities may reside in memory. Hence we index the logical
+relation over worlds, which are used to model the memory, and our logical
+relation is thus an example of what is often referred to as a Kripke
+logical relation. A world consists of a finite set of regions, and each region
 can be understood as a protocol describing the possible evolution of
@@
   interpretation function.}
-% Intuition worlds
-The worlds are step-indexed Kripke worlds. For the reader not familiar
-with these, it is sufficient to think of worlds as finite sets of
-regions that impose protocols on the memory.
 
-% Regions intuition
-We have three types of regions: $\perma$, $\temp$, and $\revoked$. The
-$\perma$ and $\temp$ regions contains a state transition system with
-public and private transitions as known from logical relations
-from high-level languages\todo{reference}. The state transition
-systems can be understood as protocols the memory must follow. The
-protocols imposed on memory by $\perma$ regions stay in place
-indefinitely. This means that any capability, $\local$ or $\glob$, can
-depend on these protocols. The protocols imposed on memory by $\temp$
-regions can be revoked at certain points during execution
-corresponding to the points when $\local$ capabilities can be
-revoked. This means that $\local$ capabilties safely can depend on the
-protocols imposed on memory by $\temp$ regions, but $\glob$
-capabilities cannot as the $\glob$ capability may stay around
-throughout the execution of a program, but the $\temp$ regions may not
-stay around.
+A world is a finite map from region names, modeled as natural numbers, to regions.
+We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
+%$\perma$, $\temp$, and $\revoked$. The
+%$\perma$ and $\temp$ 
+Each permanent and temporary region contains a state transition system, with
+public and private transitions, which describes the protocol governing
+the evolution of the memory described by the region, similarly to what
+has been used in logical relations for high-level
+languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
+The protocols imposed on memory by permanent regions stay in place
+indefinitely. This means that any capability, local or global, can
+depend on these protocols. The protocols imposed on memory by temporary
+regions can be revoked at certain points during execution,
+corresponding to the points when local capabilities can be
+revoked. This means that local capabilities safely can depend on the
+protocols imposed on memory by temporary regions, wheareas global
+capabilities cannot, since a global capability may stay around
+throughout the execution of a program, but a temporary region may be
+revoked.
 
-For monotonicity reasons, a region cannot disappear, but
-we need a way to remove $\temp$ regions. This is where the $\revoked$ region comes
-into the picture. The $\revoked$ region contains no state transition
+For technical reasons (monotonicity of the logical relation), a
+temporary region is not removed in the model when it is
+revoked. Instead it is turned into a new kind of region, namely a
+revoked region. A revoked region contains no state transition
 system and puts no requirements on the memory it models. It serves
-simply as a mask for the $\temp$ regions that have been revoked.
+simply as a mask for a temporary region that have been revoked.
+(This way of masking a region goes back to earlier work of
+Ahmed~\cite{Ahmed2004semantics} and was also used in~\cite{Thamsborg:2011:KLR:2034773.2034831}.)
 

### 2017-02-07 csf-2017/paper.tex
lr in progress

--- 2017-02-07 csf-2017/paper.tex
@@
 \begin{mathpar}
-  \inferrule{  (s,s') \in \phi \\
-    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
+              (s,s') \in \phi} 
   {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
@@
 \begin{mathpar}
-  \inferrule{  (s,s') \in \phi_\pub \\
-    (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
+    (s,s') \in \phi_\pub }
   {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
@@
 
+\todo[inline]{Put paragraph about memory satisfaction here}
+\begin{align*}
+  &\memSat{\ms}{W}
+    \text{ iff }
+    \left\{\begin{aligned}
+        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
+        &\quad (\hs = \biguplus_{r\in\activeReg{W}}P(r) \land \\
+        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
+        &\qquad W(r) = (\_,s,\_,\_,H) \land \\
+        &\qquad \npair[n]{P(r)} \in H(s)(\xi^{-1}(W)))\\
+      \end{aligned}\right.
+\end{align*}
+
 \subsection{Logical relation}
+
+\subsubsection{Observation relation}
+\todo[inline]{write something about single orthogonal closure?}  The
+\emph{observation relation}, $\observations$, is a world-indexed
+predicate that captures all the configurations that produce an
+admissable result. Here we take the result of a computation to be the
+memory when the machine halts successfully. We expect the memory to
+adhere to the protocols specified by the world, but the memory may
+change during the execution, so we just want the resulting memory to
+satisfy some private future world.
+
+We also consider a failing execution admissable, because the assembly
+language contains a $\fail$ instruction, and we want to be able to
+reason about abritrary adverserial programs.
+
+The observation relation is defined as follows:
+
+\begin{align*}
+  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
+  \observations (W) \defeq & \lambda W \ldotp 
+                             \{ \npair{(\reg,\ms)} \mid
+                             \begin{aligned}[t]
+                               & \forall \ms_f, \heap', i \leq n \ldotp \\
+                               & \quad(\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
+                               & \qquad \Rightarrow
+                               \begin{aligned}[t]
+                                 & \exists W' \futurestr W \ldotp\exists \ms_r, \ms' \ldotp\\
+                                 & \quad \heap' = \ms' \uplus \ms_r \uplus \ms_f \land \\ 
+                                 & \quad \heapSat[\hs']{n-i}{W'} \}
+                               \end{aligned}
+                             \end{aligned}
+\end{align*}
+
+\subsubsection{Expression and register-file relation}
+The expression relation, $\stder$, is a world-indexed predicate that
+contains all the words that safely can be put in the $\pcreg$-register
+of a register-file that only gives access to capabilities that are capability
+safe. 
+\begin{align*}
+  \stder : & \Worlds \nefun \UPred{\Words} \\
+  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
+                  \begin{aligned}[t]
+                    & \forall n' \leq n \ldotp\\
+                    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
+                    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
+                    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
+                  \end{aligned}
+\end{align*}
+
+The register-file relation, $\stdrr$, contains all the register files
+that contains capability safe words. It does not require the word in
+the $\pcreg$-register to be capability safe, because the relation is
+only used in the expression relation, where the $\pcreg$-register is
+overwritten anyway.
+
+\begin{align*}
+  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
+  \stdrr \defeq & \lambda W \ldotp
+                  \begin{aligned}[t]
+                    \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                    & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
+                  \end{aligned}
+\end{align*}
+
+\subsubsection{Value relation}
+The value relation, $\stdvr$, contains all the words that are safe to
+give away to an adversary.
+
+\begin{align*}
+  &\iota_{\start,\addrend}^\pwl : \Regions\\
+  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
+  \\
+  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
+  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms} \}
+\end{align*}
+Revoking all temporary regions:
+\begin{align*}
+  \revokeTemp{} & : \Worlds \fun \Worlds \\
+  \revokeTemp{W} & \defeq \lambda r \ldotp 
+                   \begin{cases}
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
+                     W(r)                & \text{otherwise}
+                   \end{cases}
+\end{align*}
+\newcommand{\wrev}[1]{\revokeTemp{#1}}
+
+\begin{align*}
+  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
+  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
+  \\
+  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \\
+      &\quad \nonlocal{\ms(\addr)} \land\\ 
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms} \}
+\end{align*}
+
+\begin{align*}
+&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
+&\qquad\begin{multlined}
+  \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
+  \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
+  \dom(\ms) = \dom(\ms') \wedge \\
+  \forall \addr \in
+  \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
+\end{multlined}
+\end{align*}
+
+\begin{align*}
+  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+  & \writeCond{}(\iota,\gl)(W) =  \\
+  & \quad \begin{aligned}[t]
+    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{ and }\\
+    & \;\qquad W(r) \text{ is address-stratified } \}
+  \end{aligned}
+\end{align*}
+
+\begin{align*}
+  & \readCond{} : \Globals \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+  & \readCond{}(\gl)(W) =  \\
+  & \quad \begin{aligned}[t]
+    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+    & \;\qquad W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl \}
+  \end{aligned}
+\end{align*}
+
+\begin{align*}
+  & \execCond{}(\gl)(W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad\forall a \in [\start,\addrend] \ldotp\\
+      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+\begin{align*}
+  & \entryCond{}(\gl)(W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+Now define the value relation as follows:
+\begin{align*}
+  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
+  &\stdvr\defeq \lambda \; W \ldotp 
+    \begin{aligned}[t]
+      & \{ \npair{i} \mid i \in \ints \} 
+      \union \\
+      & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
+      \union \\
+      & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
+      \union \\
+      & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
+      \union \\
+      & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
+      \union \\
+      & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
+      \union \\
+      & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
+      & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
+      \union \\
+      & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
+      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
+      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
+      \union \\
+      & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
+      & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
+      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
+      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
+    \end{aligned}
+\end{align*}
 

### 2017-02-07 csf-2017/paper.tex

@@
   interpretation function.}
-% Intuition worlds
-The worlds are step-indexed Kripke worlds. For the reader not familiar
-with these, it is sufficient to think of worlds as finite sets of
-regions that impose protocols on the memory.
-
-% Regions intuition
-We have three types of regions: $\perma$, $\temp$, and $\revoked$. The
-$\perma$ and $\temp$ regions contains a state transition system with
-public and private transitions as known from logical relations
-from high-level languages\todo{reference}. The state transition
-systems can be understood as protocols the memory must follow. The
-protocols imposed on memory by $\perma$ regions stay in place
-indefinitely. This means that any capability, $\local$ or $\glob$, can
-depend on these protocols. The protocols imposed on memory by $\temp$
-regions can be revoked at certain points during execution
-corresponding to the points when $\local$ capabilities can be
-revoked. This means that $\local$ capabilties safely can depend on the
-protocols imposed on memory by $\temp$ regions, but $\glob$
-capabilities cannot as the $\glob$ capability may stay around
-throughout the execution of a program, but the $\temp$ regions may not
-stay around.
-
-For monotonicity reasons, a region cannot disappear, but
-we need a way to remove $\temp$ regions. This is where the $\revoked$ region comes
-into the picture. The $\revoked$ region contains no state transition
+
+A world is a finite map from region names, modeled as natural numbers, to regions.
+We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
+%$\perma$, $\temp$, and $\revoked$. The
+%$\perma$ and $\temp$ 
+Each permanent and temporary region contains a state transition system, with
+public and private transitions, which describes the protocol governing
+the evolution of the memory described by the region, similarly to what
+has been used in logical relations for high-level
+languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
+The protocols imposed on memory by permanent regions stay in place
+indefinitely. This means that any capability, local or global, can
+depend on these protocols. The protocols imposed on memory by temporary
+regions can be revoked at certain points during execution,
+corresponding to the points when local capabilities can be
+revoked. This means that local capabilities safely can depend on the
+protocols imposed on memory by temporary regions, wheareas global
+capabilities cannot, since a global capability may stay around
+throughout the execution of a program, but a temporary region may be
+revoked.
+
+For technical reasons (monotonicity of the logical relation), a
+temporary region is not removed in the model when it is
+revoked. Instead it is turned into a new kind of region, namely a
+revoked region. A revoked region contains no state transition
 system and puts no requirements on the memory it models. It serves
-simply as a mask for the $\temp$ regions that have been revoked.
+simply as a mask for a temporary region that have been revoked.
+(This way of masking a region goes back to earlier work of
+Ahmed~\cite{Ahmed2004semantics} and was also used in~\cite{Thamsborg:2011:KLR:2034773.2034831}.)
 

### 2017-02-07 csf-2017/paper.tex
lr editing

--- 2017-02-07 csf-2017/paper.tex
@@
 revoked region. A revoked region contains no state transition
-system and puts no requirements on the memory it models. It serves
-simply as a mask for a temporary region that have been revoked.
-(This way of masking a region goes back to earlier work of
-Ahmed~\cite{Ahmed2004semantics} and was also used in~\cite{Thamsborg:2011:KLR:2034773.2034831}.)
+system and puts no requirements on the memory it models. It
+simply serves as a mask for a temporary region that have been revoked.
+This way of masking a region goes back to earlier work of
+Ahmed~\cite{Ahmed2004semantics} and was also used in~\cite{Thamsborg:2011:KLR:2034773.2034831}.
 
 % Recursive worlds intuition
-The worlds end up being recursive. The memory can contain capabilities
-and capabilities grants authority over part of the memory. A world
-models memory, so it also needs to be able to model capabilities in
-memory and the protocols on them. Whether a capability satisfies the
-protocol of a region may depend on what it grants authority over in
-memory which is described by the world.
+% The worlds end up being defined recursively: The memory can contain
+% capabilities,
+% which grant authority over parts of memory. A world
+% models memory, so it also needs to be able to model capabilities in
+% memory and the protocols on them. Whether a capability satisfies the
+% protocol of a region may depend on what it grants authority over in
+% memory which is described by the world.
+
+The worlds end up being defined recursively: Worlds describe memory,
+which can contain capabilities, whose safety depends on worlds. 
+We use the method of Birkedal et
+al.~\cite{Birkedal:2011:SKM:1926385.1926401} 
+for constructing the worlds.
 
 % Short technical explanation
-The recursiveness of the worlds gives rise to a recurisve domain
-equation. Theorem~\ref{thm:world-existence} gives the solution to this
-recursive domain equation.
+% The recursiveness of the worlds gives rise to a recurisve domain
+% equation. Theorem~\ref{thm:world-existence} gives the solution to this
+% recursive domain equation.
 \begin{theorem}\label{thm:world-existence}

### 2017-02-07 csf-2017/paper.tex
lr editing

@@
 \subsubsection{Future worlds}
-In order to model changes to memory over time, we use the future world
-relations. The future world relations model how more memory may be
-allocated but also how memory may change. How memory is allowed to
-change over time depends on the region that governs it and the future
-world relation used. The most interesting part about the future worlds
-is how it affects the regions, so we first look at the properties of
-the \emph{public future region} relation and the \emph{private future
-  region} relation. We then use these relations to describe the future
-world relations.
-
-{\bf Private future regions}
-In the private future region relation, we allow revocation of $\temp$
-regions. The regions cannot be revoked by removing them from the
-world, because we need the future world relations to be
-monotone. This is why we have the $\revoked$ region, so we have
-a region that can serve as a mask. We can, however, let any region
-serve as a mask, so in the private future region relation, we allow a
-$\temp$ region to become any valid region. As the $\revoked$ region
-serves as a mask, we let any valid region take its place in the
-private future region relation.
-
-If a $\temp$ region is not masked by a region, then it behaved like a
-$\perm$ region. The $\perma$ regions stay the same type and they are
-allowed to take a transition in the private part of the transition
-system. Which can be thought of as the memory changes according to
-some protocol.
+The future world relations model how memory may evolve over time. 
+The \emph{public future world} 
+$W' \futurewk W$ is defined to mean that $\dom(W') \supseteq \dom(W)$
+and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$.  That is,
+in a public future world, more regions may have been allocated, and
+each of the regions may also have evolved according to the public future
+region relation (defined below). The \emph{private future world} relation
+$W' \futurestr W$ is defined similarly, using a private future region
+relation.
 
@@
 \end{mathpar}
-
-{\bf Public future regions}
-In the public future regions we do not permit revocation of $\temp$
-regions. In this relation, the $\temp$ and $\perma$ regions must stay
-the same type and they are allowed to take a transition in the public
-part of the transition system.
-
-The $\revoked$ regions serve as masking regions, but are not as
-liberal as in the private future region relation when it comes to
-letting other regions replace them as masks. In the public future
-region relation, we only allow $\revoked$ and $\temp$ regions to take
-on the job as masking region. We have this limitation, because we
-want the private future region relation to be able to reuse any region
-it has revoked.
+These rules express that, in the private future region relation, 
+we allow revocation of temporary regions. 
+As explained above, this can be done using a special revoked region,
+which does not desribe any memory, but we also allow  
+any other region to serve as a mask, and thus any region $r$ can be a
+future region of a temporary region and of a revoked region.
+If a $\temp$ region is not masked by a region, then it behaves like a
+$\perm$ region. The $\perma$ regions stay permanent and they are
+allowed to take a transition in the private part of the transition
+system.
 
@@
 \end{mathpar}
+In the public future regions we do not permit revocation of temporary
+regions. Instead, temporary and permanent regions are only 
+allowed to take a transition in the public part of the transition system.
+
+The $\revoked$ regions serve as masking regions, but are not as
+liberal as in the private future region relation when it comes to
+letting other regions replace them as masks. In the public future
+region relation, we only allow $\revoked$ and $\temp$ regions to take
+on the job as masking region. We have this limitation, because we
+want the private future region relation to be able to reuse any region
+it has revoked.
+\todo[inline]{can we make this crisper ?}
 
 Notice that the public future region relation is a subset of the
-private future region relation, so anything you are allowed to do in
-the public future region relation, the private future region relation
-can do the same.
+private future region relation.
 
-% The future world relation are really the ones from thm, the above
-% are properties of them.
-{\bf Future world relations}
-There are two future world relations a \emph{public future world}
-relation and a \emph{private future world} relation. In both future
-world relations, we have extension ordering, so a future world $W'$ of
-$W$ has at least the same regions as $W$, but $W'$ may have new regions
-imposing new protocols on parts of the memory. The two future world
-relations differ in the way existing regions are allowed to
-evolve. The public future world relation uses the public future region
-relation and the private future world relation uses the private future
-region relation:
-\begin{mathpar}
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-    \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
-  { W' \futurewk W }
-  \and
-  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-    \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
-  { W' \futurestr W }
-\end{mathpar}
+%%%%%%%%
 
-% Relate to STSs for high-level languages. Have the public-private
-% transitions, but they are also used for more than this
-\todo[inline]{Relate to STSs for high-level languages?}
+% In order to model changes to memory over time, we use the future world
+% relations. The future world relations model how more memory may be
+% allocated but also how memory may change. How memory is allowed to
+% change over time depends on the region that governs it and the future
+% world relation used. The most interesting part about the future worlds
+% is how it affects the regions, so we first look at the properties of
+% the \emph{public future region} relation and the \emph{private future
+%   region} relation. We then use these relations to describe the future
+% world relations.
+
+% {\bf Private future regions}
+% In the private future region relation, we allow revocation of $\temp$
+% regions. The regions cannot be revoked by removing them from the
+% world, because we need the future world relations to be
+% monotone. This is why we have the $\revoked$ region, so we have
+% a region that can serve as a mask. We can, however, let any region
+% serve as a mask, so in the private future region relation, we allow a
+% $\temp$ region to become any valid region. As the $\revoked$ region
+% serves as a mask, we let any valid region take its place in the
+% private future region relation.
+
+% If a $\temp$ region is not masked by a region, then it behaved like a
+% $\perm$ region. The $\perma$ regions stay the same type and they are
+% allowed to take a transition in the private part of the transition
+% system. Which can be thought of as the memory changes according to
+% some protocol.
+
+% The properties of the \emph{private future region} relation are:
+% \begin{mathpar}
+%   \inferrule{  (s,s') \in \phi \\
+%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+%   {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
+%   \and
+%   \inferrule{ r \in \Regions }
+%   { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
+%   \and
+%   \inferrule{ r \in \Regions }
+%   { r \futurestr \revoked }
+% \end{mathpar}
+
+% {\bf Public future regions}
+% In the public future regions we do not permit revocation of $\temp$
+% regions. In this relation, the $\temp$ and $\perma$ regions must stay
+% the same type and they are allowed to take a transition in the public
+% part of the transition system.
+
+% The $\revoked$ regions serve as masking regions, but are not as
+% liberal as in the private future region relation when it comes to
+% letting other regions replace them as masks. In the public future
+% region relation, we only allow $\revoked$ and $\temp$ regions to take
+% on the job as masking region. We have this limitation, because we
+% want the private future region relation to be able to reuse any region
+% it has revoked.
+
+% The \emph{public future} region relations satisfies the following properties:
+% \begin{mathpar}
+%   \inferrule{  (s,s') \in \phi_\pub \\
+%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
+%   {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
+%   \and
+%   \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
+%   { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
+%   \and
+%   \inferrule{ }
+%   { \revoked \futurewk \revoked }
+% \end{mathpar}
+
+% Notice that the public future region relation is a subset of the
+% private future region relation, so anything you are allowed to do in
+% the public future region relation, the private future region relation
+% can do the same.
+
+% % The future world relation are really the ones from thm, the above
+% % are properties of them.
+% {\bf Future world relations}
+% There are two future world relations a \emph{public future world}
+% relation and a \emph{private future world} relation. In both future
+% world relations, we have extension ordering, so a future world $W'$ of
+% $W$ has at least the same regions as $W$, but $W'$ may have new regions
+% imposing new protocols on parts of the memory. The two future world
+% relations differ in the way existing regions are allowed to
+% evolve. The public future world relation uses the public future region
+% relation and the private future world relation uses the private future
+% region relation:
+% \begin{mathpar}
+%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+%     \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
+%   { W' \futurewk W }
+%   \and
+%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
+%     \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
+%   { W' \futurestr W }
+% \end{mathpar}
+
+% % Relate to STSs for high-level languages. Have the public-private
+% % transitions, but they are also used for more than this
+% \todo[inline]{Relate to STSs for high-level languages?}
 

### 2017-02-07 csf-2017/paper.tex
A pass over the introduction.

@@
 \usepackage{hyperref}
+\usepackage{natbib}
 \lstset{
@@
 \todo[inline]{See the latex comments for the next paragraphs!}
-% Possible high-level intros:
-% 0) Arms race compilers
-%
-% 1) Security guarantees vs obstacles / arms race
-\begin{comment}
-  Security on processors is today subject to an arms race between
-  processor manufacturers and hackers. On the hacker side, a constant
-  effort is put into exploiting current processor designs and
-  circumventing the security measures set in place to prevent known
-  exploits. On the processor manufacturer side, new security measures
-  are put in place to prevent new exploits. The arms race would come
-  to a grinding halt if the security meassures provided more than
-  obstacles and actually provided low-level security guarantees. One
-  kind of low-level machines that provide low-level security is a
-  capability machine.
-\end{comment}
-% 2) Fully abstract compilation
-High-level languages provide guarantees such as encapsulation of local
-state and well-bracketedness of function calls. These guarantees are
-often relied upon when reasoning about the correctness and security of
-a program. High-level programs are compiled to machine code, so in
-order to rely on the correctness and security guarantees provided by
-the high-level language, it needs to be proven that the translation
-preserves these properties. Translated programs often interact with
-machine code that is not translated from the same high-level language,
-so in order to preserve the guarantees from the high-level languages,
-the low-level machine needs to provide some kind of security
-guarantees. In other words, a target of secure compilation needs to
-provide some security guarantees. Modern processors provide many
-security meassures, but very little in terms of guarantees. A machine
-that provides some means of security guarantees is a capability machine.
-
-% 3) Memory access-control
-% I don't like this angle and can't find a good way to sell it.
-
-% 4) History
-% Proposals throughout history.
-% Lack of formal account of properties-guarantees.
-
-%%% a)
-% Current low-level protection coarse grained memory protection or
-% properties of high-level languages are not really enforced?
-
-% Capability machines offer fine-grained memory protection
-% Too technical
-A capability machine is a low-level machine that offers fine-grained
-memory protection.
-% Capability machines and unforgeable tokens for memory access. 
-This is done by introducing unforgeable tokens to
-the machine. These tokens are called capabillities and grant a kind of
-authority and a range of authority. 
-% Dynamic checks encapsulation, formal model
-
- We will refer to the capabilities
-that grant a combination of read, write, or execute permission as memory
-capabilities. 
- Capabilities with the necessary permissions must be
-presented every time an instruction that manipulates the memory is
-executed.
-
-Memory capabilities provide memory protection, but they do not provide
-a way to set up security domains. When you jump to an untrusted piece
-of code that is supposed to jump back to you, then you have to either
-revoke all your capabilities or pass to the piece of code you do not
-trust. This short coming can be overcome by adding something like the
-enter capability from the M-Machine\todo{add reference}. The enter
-capability is an opaque capability which can only be used for a
-jump. When jumped to it grants read and execute permission which
-allows the code that is now executing to read capabilities stored in
-memory. 
-The CHERI processor's ccall achieves a similar property\todo{Add reference}.
-
-Capabilities are irrevocable, so when we pass a capability has to
-an untrusted program, then we have to assume that the program keeps it
-around indefenitely. This means that we cannot reuse the piece of
-memory that the capability governs essentially creating a memory
-leak. The CHERI processor has a special kind of capabilities called
-\emph{local capabilities}\todo{reference?}\ that introduces a simple
-kind of temporal information control which allows for a simple kind of
-capability revocation. This is achieved by adding a tag to every
-capability that marks it as either local or global. Local capabilities
-can only be written through a capability with a new permission called
-\emph{permit write local}.
-% More about local capabilities.
-
-With just memory capabilities, enter capabailities, and local
-capabilities, the capability machine is very powerful and can
-enforce properties that high-level languages promise. In most
-high-level languages, we expect local state to be private and
-encapsulated. The memory capabilities make sure that local state
-cannot be accessed without a capability to do so. With the enter
-capabaility, we can regain control of local state after passing
-control to an untrusted piece of code which gives us encapsulation
-when control is passed to an untrusted program.
-
-Many high-level languages guarantee that function function calls are
-well-bracketed. This can easily be achieved by having a trusted call
-stack. On a capability machine, it is possible to enforce well
-barcketedness without a trusted stack. When ensuring well-bracketed
-the two main challenges calls are 1) preventing storage of return
-pointers and 2) encapsulation of the local stack frame. If we don't
-have the first property, then an adversary can store the return
-capability in a call and jump to it in a later call. If the local
-stack frame is not properly encapsulated, then an adversary can break
-the well-bracketedness by jumping to a different calls return
-capability.
-
-In order to deal with the two challenges, we make sure that all return
-capabilities are local and that there are no global capabilities with
-permit write local permission. This makes sure that there is no way to
-save a return capability in persistent local state.\lau{this does not really make
-sense as the local stack frame can be seen as local state} It is,
-however, to limiting to not provide any means to store the return
-capabilities as a piece of code may need to store a return capability
-before jumping to an adversary. It is therefore needed to provide a
-local capability with permit write local permission which can be used
-as a stack. The stack should now be the only place one can store the
-return capability, but it should also be the only place one can store
-the stack capability.
-
-Using local return capabilities and putting a stack abstraction on a
-local capability permit write local capability is a good start, but as
-it turns out, it is not wnough. An adversary could do the following 
-\begin{enumerate}
-\item In a first invocation, fill the stack frame with copies of the
-  current stack capability.
-\item In a later call, the adversary may be able to load the stack
-  capability from the previous call.
-\item If we have used a larger part of the stack than in the first
-  call, then the adversary has access to part of our local stack frame.
-\end{enumerate}
-
-The above attack shows that it is tricky to ensure
-well-bracketedness. It raises the question whether it is possible to
-do so it is completely watertight. The attack utilized the stack, so
-it is necessary to have a trusted stack or can we do without? In
-general it raises the question, how do we reason about capability
-machines and local capabilities, so we are sure that what ever calling
-scheme we come up with actually provably ensures well-bracketed calls.
-
-In the following paper, we present the following contributions
-\begin{enumerate}
-\item 
-\end{enumerate}
+
+Compromising software security is often based on attacks that break programming
+language properties relied on by software authors, such as control flow
+integrity, local state encapsulation etc. Commodity processors offer little
+support for defending against such attacks: they offer security primitives with
+only coarse-grained memory protection and limited compartmentalization
+scalability. As a result, defenses against attacks on control flow integrity and
+local state encapsulation necessarily rely on techniques like machine code
+rewriting (e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}) or
+verification (e.g.~\cite{morrisett_system_1999}), virtual machines with a native
+stack (e.g.~the JVM~\cite{lindholm_java_2014}) or
+randomization~\cite{forrest_building_1997}. Such techniques for retrofitting
+protection techniques to commodity hardware have disadvantages in terms of
+performance, complexity and/or security.
+
+\emph{Capability machines} are a type of processors, intended to remediate these
+limitations by presenting a better security model at the hardware
+level~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740,Carter:1994:HSF:195473.195579}.\dominique{todo:
+  more?} Such hardware tags every word (in the register file and in memory) to
+enforce a strict separation between numbers and capabilities, a kind of pointers
+carrying some kind of authority. Memory capabilities carry the authority to read
+and/or write to a range of memory locations. Additionally, the details vary but
+all capability machines offer some form of \emph{object capabilities}, which
+represent the authority to invoke a piece of code without exposing the code's
+encapsulated private state (e.g. the M-Machine's enter capabilities or CHERI's
+sealed code/data pairs).
+
+Unlike for commodity processors, the security primitives on capability machines
+lend themselves well to the enforcement of properties like local state
+encapsulation and control-flow integrity. Potentially, they will enable new
+compilation schemes that enforce such properties in a way that is efficient but
+also 100\% watertight (ideally evidenced by a mathematical proof, guaranteeing
+that we do not end up in a new attack-defense arms race). However, there is
+still a lot that needs to happen to attain this goal. For example, it is quite
+non-trivial to devise a compilation scheme adapted to the details of a specific
+language's notion of encapsulation (e.g. encapsulation of private member
+variables in OO languages often behaves quite differently than encapsulation of
+private state in ML-like languages). And even if such a scheme were defined, a
+formal proof depends on a formalisation of the encapsulation provided by the
+capability machine at hand.
+
+A similar problem is the enforcement of control-flow integrity on capability
+machines. An interesting approach is taken in CheriBSD. They drop the standard
+contiguous C stack and split it into a central, trusted shadow stack, managed by
+trusted call and return instructions\footnote{These call and return instructions
+  are implemented by CheriBSD and special support from the CHERI processor makes
+  them very fast to invoke.}, and disjoint, per-component
+stacks~\citep{Watson2015Cheri}. For preventing illegal use of stack references,
+the approach relies on \emph{local capabilities}, a type of capabilities offered
+by CHERI that can be passed over a function invocation \emph{temporarily}, for
+the duration of the invocation, and revoked afterwards. However, a lot of
+details remain to be filled in (e.g. how to combine a trusted shadow stack with
+(potentially) cross-domain function pointers in a scalable way?) and there is no
+argument that this enforcement of CFI is watertight.
+
+In this paper, we make progress towards the goal of efficient, provably
+watertight enforcement of encapsulation and control-flow integrity on capability
+machines.  Specifically, we make the following contributions:
+\begin{itemize}
+\item We formalise a simple but representative capability machine featuring
+  local capabilities much like CHERI's, and define its operational semantics.
+\item We define a novel stack management discipline for enforcing control-flow
+  integrity and encapsulation of local state on the stack. The discipline relies
+  solely on local capabilities (used to represent stack and return pointers) and
+  does not require OS support in the form of a trusted shadow stack or special
+  call and return instructions. It also does not require separate per-component
+  stacks and instead works with a single, fairly standard contiguous stack. It
+  supports higher-order cross-component calls (e.g. (potentially)
+  cross-component function pointers) and can be efficient assuming one
+  additional piece of processor support: an instruction for efficiently clearing
+  a range of memory.
+\item We provide a theorem (capability-safety) that is (by far) the most general
+  and powerful formulation of the guarantees offered by a capability machine.
+  The theorem also expresses the specific guarantees offered by the machine for
+  local capabilities. It is very general and not tied to a specific ABI or a
+  specific way of using the system's capabilities.
+\item The above theorem is formulated in terms of a unary step-indexed logical
+  relation with recursively defined Kripke worlds. It is inspired by one used by
+  \cite{Devriese:2016ObjCap} for capability-safety in a JavaScript-like lambda
+  calculus, but has two novel technical characteristics. The natural CPS style
+  of assembly code with return pointers as continuations is modeled using a
+  \emph{single} orthogonal closure, rather than the biorthogonal closure that
+  has been previously used for lambda calculi. Perhaps more surprisingly, we
+  show that the special nature of local capabilities can be elegantly expressed
+  using a variant of an idea that has been used by
+  \cite{Dreyer:2010:IHS:1863543.1863566} to express well-bracketedness
+  guarantees in a lambda calculus: Kripke worlds with a weak and a strong future
+  world relations. Interestingly, the two future world relations generally work
+  very similarly to Dreyer et al.'s, except that they turn up in quite a
+  different place in the logical relation.
+\item We demonstrate the power of our results by applying them to challenging
+  examples, specifically constructed to demonstrate local state encapsulation
+  and control-flow integrity guarantees in the presence of cross-component
+  function pointers. The examples demonstrate both the power of our formulation
+  of capability-safety and our stack management discipline.
+\end{itemize}
+
+% % Possible high-level intros:
+% % 0) Arms race compilers
+% %
+% % 1) Security guarantees vs obstacles / arms race
+% \begin{comment}
+%   Security on processors is today subject to an arms race between
+%   processor manufacturers and hackers. On the hacker side, a constant
+%   effort is put into exploiting current processor designs and
+%   circumventing the security measures set in place to prevent known
+%   exploits. On the processor manufacturer side, new security measures
+%   are put in place to prevent new exploits. The arms race would come
+%   to a grinding halt if the security meassures provided more than
+%   obstacles and actually provided low-level security guarantees. One
+%   kind of low-level machines that provide low-level security is a
+%   capability machine.
+% \end{comment}
+% % 2) Fully abstract compilation
+% High-level languages provide guarantees such as encapsulation of local
+% state and well-bracketedness of function calls. These guarantees are
+% often relied upon when reasoning about the correctness and security of
+% a program. High-level programs are compiled to machine code, so in
+% order to rely on the correctness and security guarantees provided by
+% the high-level language, it needs to be proven that the translation
+% preserves these properties. Translated programs often interact with
+% machine code that is not translated from the same high-level language,
+% so in order to preserve the guarantees from the high-level languages,
+% the low-level machine needs to provide some kind of security
+% guarantees. In other words, a target of secure compilation needs to
+% provide some security guarantees. Modern processors provide many
+% security meassures, but very little in terms of guarantees. A machine
+% that provides some means of security guarantees is a capability machine.
+
+% % 3) Memory access-control
+% % I don't like this angle and can't find a good way to sell it.
+
+% % 4) History
+% % Proposals throughout history.
+% % Lack of formal account of properties-guarantees.
+
+% %%% a)
+% % Current low-level protection coarse grained memory protection or
+% % properties of high-level languages are not really enforced?
+
+% % Capability machines offer fine-grained memory protection
+% % Too technical
+% A capability machine is a low-level machine that offers fine-grained
+% memory protection.
+% % Capability machines and unforgeable tokens for memory access. 
+% This is done by introducing unforgeable tokens to
+% the machine. These tokens are called capabillities and grant a kind of
+% authority and a range of authority. 
+% % Dynamic checks encapsulation, formal model
+
+%  We will refer to the capabilities
+% that grant a combination of read, write, or execute permission as memory
+% capabilities. 
+%  Capabilities with the necessary permissions must be
+% presented every time an instruction that manipulates the memory is
+% executed.
+
+% Memory capabilities provide memory protection, but they do not provide
+% a way to set up security domains. When you jump to an untrusted piece
+% of code that is supposed to jump back to you, then you have to either
+% revoke all your capabilities or pass to the piece of code you do not
+% trust. This short coming can be overcome by adding something like the
+% enter capability from the M-Machine\todo{add reference}. The enter
+% capability is an opaque capability which can only be used for a
+% jump. When jumped to it grants read and execute permission which
+% allows the code that is now executing to read capabilities stored in
+% memory. 
+% The CHERI processor's ccall achieves a similar property\todo{Add reference}.
+
+% Capabilities are irrevocable, so when we pass a capability has to
+% an untrusted program, then we have to assume that the program keeps it
+% around indefenitely. This means that we cannot reuse the piece of
+% memory that the capability governs essentially creating a memory
+% leak. The CHERI processor has a special kind of capabilities called
+% \emph{local capabilities}\todo{reference?}\ that introduces a simple
+% kind of temporal information control which allows for a simple kind of
+% capability revocation. This is achieved by adding a tag to every
+% capability that marks it as either local or global. Local capabilities
+% can only be written through a capability with a new permission called
+% \emph{permit write local}.
+% % More about local capabilities.
+
+% With just memory capabilities, enter capabailities, and local
+% capabilities, the capability machine is very powerful and can
+% enforce properties that high-level languages promise. In most
+% high-level languages, we expect local state to be private and
+% encapsulated. The memory capabilities make sure that local state
+% cannot be accessed without a capability to do so. With the enter
+% capabaility, we can regain control of local state after passing
+% control to an untrusted piece of code which gives us encapsulation
+% when control is passed to an untrusted program.
+
+% Many high-level languages guarantee that function function calls are
+% well-bracketed. This can easily be achieved by having a trusted call
+% stack. On a capability machine, it is possible to enforce well
+% barcketedness without a trusted stack. When ensuring well-bracketed
+% the two main challenges calls are 1) preventing storage of return
+% pointers and 2) encapsulation of the local stack frame. If we don't
+% have the first property, then an adversary can store the return
+% capability in a call and jump to it in a later call. If the local
+% stack frame is not properly encapsulated, then an adversary can break
+% the well-bracketedness by jumping to a different calls return
+% capability.
+
+% In order to deal with the two challenges, we make sure that all return
+% capabilities are local and that there are no global capabilities with
+% permit write local permission. This makes sure that there is no way to
+% save a return capability in persistent local state.\lau{this does not really make
+% sense as the local stack frame can be seen as local state} It is,
+% however, to limiting to not provide any means to store the return
+% capabilities as a piece of code may need to store a return capability
+% before jumping to an adversary. It is therefore needed to provide a
+% local capability with permit write local permission which can be used
+% as a stack. The stack should now be the only place one can store the
+% return capability, but it should also be the only place one can store
+% the stack capability.
+
+% Using local return capabilities and putting a stack abstraction on a
+% local capability permit write local capability is a good start, but as
+% it turns out, it is not wnough. An adversary could do the following 
+% \begin{enumerate}
+% \item In a first invocation, fill the stack frame with copies of the
+%   current stack capability.
+% \item In a later call, the adversary may be able to load the stack
+%   capability from the previous call.
+% \item If we have used a larger part of the stack than in the first
+%   call, then the adversary has access to part of our local stack frame.
+% \end{enumerate}
+
+% The above attack shows that it is tricky to ensure
+% well-bracketedness. It raises the question whether it is possible to
+% do so it is completely watertight. The attack utilized the stack, so
+% it is necessary to have a trusted stack or can we do without? In
+% general it raises the question, how do we reason about capability
+% machines and local capabilities, so we are sure that what ever calling
+% scheme we come up with actually provably ensures well-bracketed calls.
+
+% In the following paper, we present the following contributions
+% \begin{enumerate}
+% \item 
+% \end{enumerate}
 

### 2017-02-08 csf-2017/paper.tex
comment

--- 2017-02-07 csf-2017/paper.tex
@@
 level~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740,Carter:1994:HSF:195473.195579}.\dominique{todo:
-  more?} Such hardware tags every word (in the register file and in memory) to
+  more?} \lars{can we add that there has been renewed interest and
+  refer to crash-safe project ?}
+ Such hardware tags every word (in the register file and in memory) to
 enforce a strict separation between numbers and capabilities, a kind of pointers

### 2017-02-08 csf-2017/paper.tex
.

--- 2017-02-08 csf-2017/paper.tex
@@
 % Configuration end states
-\newcommand{\failed}{\textsl{failed}}
-\newcommand{\halted}{\textsl{halted}}
+\newcommand{\failed}{\mathit{failed}}
+\newcommand{\halted}{\mathit{halted}}
 
@@
 The value relation, $\stdvr$, contains all the words that are safe to
-give away to an adversary.
+give away to an adversary. The conditions a capability must satisfy to
+be safe to give to an adversary depends on what permissions it
+gives. We there for define four conditions
 

### 2017-02-08 csf-2017/paper.tex
lr

@@
 
+\emph{memory}  we don't make a distinction between memory and memory
+segments in the text  (to do so properly, we would need to explain why
+assumptions
+are reasonable, given that a real machine has finite memory, but we
+need infinite domain for malloc spec, etc.)
+
 \section{Introduction}
@@
 future region of a temporary region and of a revoked region.
-If a $\temp$ region is not masked by a region, then it behaves like a
-$\perm$ region. The $\perma$ regions stay permanent and they are
+If a temporary region is not masked by a region, then it behaves like a
+permanent region. The permanent regions stay permanent and they are
 allowed to take a transition in the private part of the transition
@@
 allowed to take a transition in the public part of the transition system.
-
-The $\revoked$ regions serve as masking regions, but are not as
-liberal as in the private future region relation when it comes to
-letting other regions replace them as masks. In the public future
-region relation, we only allow $\revoked$ and $\temp$ regions to take
-on the job as masking region. We have this limitation, because we
-want the private future region relation to be able to reuse any region
-it has revoked.
-\todo[inline]{can we make this crisper ?}
+In the public future region relation, we only allow revoked and
+temporary regions to mask a region. This means that the private future
+world relations allows us to reuse a region that has been revoked
+earlier.
 
@@
 
+A memory satisfies a world, written $\memSat{\ms}{W}$, 
+if it can be partitioned into disjoint parts such that each part is
+in the interpretation of an active (permanent or temporary) region.
+
 \subsection{Logical relation}
@@
 \subsubsection{Observation relation}
-\todo[inline]{write something about single orthogonal closure?}  The
-\emph{observation relation}, $\observations$, is a world-indexed
-predicate that captures all the configurations that produce an
-admissable result. Here we take the result of a computation to be the
-memory when the machine halts successfully. We expect the memory to
-adhere to the protocols specified by the world, but the memory may
-change during the execution, so we just want the resulting memory to
-satisfy some private future world.
-
-We also consider a failing execution admissable, because the assembly
-language contains a $\fail$ instruction, and we want to be able to
-reason about abritrary adverserial programs.
-
-The observation relation is defined as follows:
-
+\todo[inline]{write something about single orthogonal closure?}
+
+ The \emph{observation relation}, $\observations$, defines which
+configurations are capability safe. Roughly speaking,
+a configuration is capability safe if, whenever it halts, then
+the resulting memory satisfies a private future world. 
+Notice that if a configuation fails, e.g., if an unauthorized access 
+(e.g., a load of a memory location without the requisite read capability)
+is attempted, then the configuation is considered capability safe, which it should
+because then the capability machine has caught the attempted unauthorized access
+This is similar to the logical relation given for an untyped language
+in~\cite{Devriese:2016ObjCap}, but unlike logical relations for typed languages, which
+typically require that computations do not fail.
+
+\subsubsection{Expression and register-file relation}
+The expression relation, $\stder$, is a world-indexed predicate that
+contains all the words that safely can be put in the $\pcreg$-register
+of a register-file that only gives access to capabilities that are capability
+safe.
+
+\begin{figure}[htbp]
+  \centering
 \begin{align*}
@@
 \end{align*}
-
-\subsubsection{Expression and register-file relation}
-The expression relation, $\stder$, is a world-indexed predicate that
-contains all the words that safely can be put in the $\pcreg$-register
-of a register-file that only gives access to capabilities that are capability
-safe. 
 \begin{align*}
@@
 \end{align*}
-
-The register-file relation, $\stdrr$, contains all the register files
-that contains capability safe words. It does not require the word in
-the $\pcreg$-register to be capability safe, because the relation is
-only used in the expression relation, where the $\pcreg$-register is
-overwritten anyway.
-
-\begin{align*}
-  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
-  \stdrr \defeq & \lambda W \ldotp
-                  \begin{aligned}[t]
-                    \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
-                    & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
-                  \end{aligned}
-\end{align*}
-
-\subsubsection{Value relation}
-The value relation, $\stdvr$, contains all the words that are safe to
-give away to an adversary. The conditions a capability must satisfy to
-be safe to give to an adversary depends on what permissions it
-gives. We there for define four conditions
-
-\begin{align*}
-  &\iota_{\start,\addrend}^\pwl : \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
-  \\
-  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
-    \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
-    \end{aligned}
-        \right\} \union \{\npair[0]{\ms} \}
-\end{align*}
-Revoking all temporary regions:
-\begin{align*}
-  \revokeTemp{} & : \Worlds \fun \Worlds \\
-  \revokeTemp{W} & \defeq \lambda r \ldotp 
-                   \begin{cases}
-                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
-                     W(r)                & \text{otherwise}
-                   \end{cases}
-\end{align*}
-\newcommand{\wrev}[1]{\revokeTemp{#1}}
-
-\begin{align*}
-  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
-  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
-  \\
-  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
-    \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \nonlocal{\ms(\addr)} \land\\ 
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
-    \end{aligned}
-        \right\} \union \{\npair[0]{\ms} \}
-\end{align*}
-
-\begin{align*}
-&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
-&\qquad\begin{multlined}
-  \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
-  \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
-  \dom(\ms) = \dom(\ms') \wedge \\
-  \forall \addr \in
-  \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
-\end{multlined}
-\end{align*}
-
-\begin{align*}
-  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
-  & \writeCond{}(\iota,\gl)(W) =  \\
-  & \quad \begin{aligned}[t]
-    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{ and }\\
-    & \;\qquad W(r) \text{ is address-stratified } \}
-  \end{aligned}
-\end{align*}
-
-\begin{align*}
-  & \readCond{} : \Globals \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
-  & \readCond{}(\gl)(W) =  \\
-  & \quad \begin{aligned}[t]
-    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-    & \;\qquad W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl \}
-  \end{aligned}
-\end{align*}
-
-\begin{align*}
-  & \execCond{}(\gl)(W) = \\
-  & \quad
-    \begin{aligned}[t]
-      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
-      & \quad \forall W' \future W \ldotp\\
-      & \qquad\forall a \in [\start,\addrend] \ldotp\\
-      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
-    \end{aligned} \\
-  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
-\end{align*}
-
-\begin{align*}
-  & \entryCond{}(\gl)(W) = \\
-  & \quad
-    \begin{aligned}[t]
-      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
-      & \quad \forall W' \future W \ldotp\\
-      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
-    \end{aligned} \\
-  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
-\end{align*}
-
-Now define the value relation as follows:
 \begin{align*}
@@
 \end{align*}
+  \caption{Logical Relation}
+  \label{fig:logrel}
+\end{figure}
+
+The register-file relation, $\stdrr$, contains all the register files
+that contains capability safe words. It does not require the word in
+the $\pcreg$-register to be capability safe, because the relation is
+only used in the expression relation, where the $\pcreg$-register is
+overwritten anyway.
+
+\begin{align*}
+  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
+  \stdrr \defeq & \lambda W \ldotp
+                  \begin{aligned}[t]
+                    \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                    & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
+                  \end{aligned}
+\end{align*}
+
+\subsubsection{Value relation}
+The value relation, $\stdvr$, contains all the words that are safe to
+give away to an adversary. The conditions a capability must satisfy to
+be safe to give to an adversary depends on what permissions it
+gives. We there for define four conditions
+
+\begin{align*}
+  &\iota_{\start,\addrend}^\pwl : \Regions\\
+  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
+  \\
+  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
+  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms} \}
+\end{align*}
+Revoking all temporary regions:
+\begin{align*}
+  \revokeTemp{} & : \Worlds \fun \Worlds \\
+  \revokeTemp{W} & \defeq \lambda r \ldotp 
+                   \begin{cases}
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
+                     W(r)                & \text{otherwise}
+                   \end{cases}
+\end{align*}
+\newcommand{\wrev}[1]{\revokeTemp{#1}}
+
+\begin{align*}
+  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
+  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
+  \\
+  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp \\
+      &\quad \nonlocal{\ms(\addr)} \land\\ 
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms} \}
+\end{align*}
+
+\begin{align*}
+&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
+&\qquad\begin{multlined}
+  \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
+  \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
+  \dom(\ms) = \dom(\ms') \wedge \\
+  \forall \addr \in
+  \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
+\end{multlined}
+\end{align*}
+
+\begin{align*}
+  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+  & \writeCond{}(\iota,\gl)(W) =  \\
+  & \quad \begin{aligned}[t]
+    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{ and }\\
+    & \;\qquad W(r) \text{ is address-stratified } \}
+  \end{aligned}
+\end{align*}
+
+\begin{align*}
+  & \readCond{} : \Globals \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+  & \readCond{}(\gl)(W) =  \\
+  & \quad \begin{aligned}[t]
+    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
+    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+    & \;\qquad W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl \}
+  \end{aligned}
+\end{align*}
+
+\begin{align*}
+  & \execCond{}(\gl)(W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad\forall a \in [\start,\addrend] \ldotp\\
+      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+\begin{align*}
+  & \entryCond{}(\gl)(W) = \\
+  & \quad
+    \begin{aligned}[t]
+      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
+      & \quad \forall W' \future W \ldotp\\
+      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+    \end{aligned} \\
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+\end{align*}
+
+Now define the value relation as follows:
 

### 2017-02-08 csf-2017/paper.tex
added figures with LR definitions

@@
 \begin{align*}
-  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-  \observations (W) \defeq & \lambda W \ldotp 
-                             \{ \npair{(\reg,\ms)} \mid
+  & \observations \; : \;  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
+  & \observations (W) \defeq \\
+  & \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
+    \begin{aligned}
+      & \forall \ms_f, \heap', i \leq n \ldotp \\
+      & (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
+      & \;\; \Rightarrow
       \begin{aligned}[t]
-                               & \forall \ms_f, \heap', i \leq n \ldotp \\
-                               & \quad(\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
-                               & \qquad \Rightarrow
-                               \begin{aligned}[t]
-                                 & \exists W' \futurestr W \ldotp\exists \ms_r, \ms' \ldotp\\
-                                 & \quad \heap' = \ms' \uplus \ms_r \uplus \ms_f \land \\ 
-                                 & \quad \heapSat[\hs']{n-i}{W'} \}
+        & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
+        & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \land \\ 
+        & \;\; \heapSat[\hs']{n-i}{W'} 
       \end{aligned}
-                             \end{aligned}
-\end{align*}
-\begin{align*}
-  \stder : & \Worlds \nefun \UPred{\Words} \\
-  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
-                  \begin{aligned}[t]
-                    & \forall n' \leq n \ldotp\\
-                    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
-                    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
-                    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
-                  \end{aligned}
-\end{align*}
-\begin{align*}
-  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
-  &\stdvr\defeq \lambda \; W \ldotp 
-    \begin{aligned}[t]
+    \end{aligned}\right\}\\
+  & \stdrr \; : \; \Worlds \monwknefun \UPred{\Regs} \\
+  & \stdrr(W) \defeq \left\{ \npair{\reg} \; \middle| \;
+                  \begin{aligned}
+                      & \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
+                    & \;\;  \npair{\reg(r)} \in \stdvr(W) 
+                  \end{aligned} \right\} \\
+  & \\
+  & \stder \; : \; \Worlds \nefun \UPred{\Words} \\
+  & \stder(W) \defeq \left\{ \npair{\pc} \; \middle| \; 
+                  \begin{aligned}
+                    & \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
+                    & \;\;\npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) 
+                  \end{aligned} \right\} \\
+  & \\
+  &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
+  &\stdvr(W)\defeq  \\
+    &\;\;\begin{aligned}[t]
       & \{ \npair{i} \mid i \in \ints \} 
@@
       & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
+      & \quad \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \} 
       \union \\
@@
       & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
-      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
-      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
+      & \quad \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) 
       \union \\
@@
       & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
-      & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
-      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
-      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
+      & \quad \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \}
     \end{aligned}
@@
 
-\begin{align*}
-  \stdrr : & \Worlds \monwknefun \UPred{\Regs} \\
-  \stdrr \defeq & \lambda W \ldotp
-                  \begin{aligned}[t]
-                    \{ \npair{\reg} \mid & \;\forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
-                    & \;\quad  \npair{\reg(r)} \in \stdvr(W) \}
-                  \end{aligned}
-\end{align*}
-
 \subsubsection{Value relation}
@@
 
+\begin{figure}[htbp]
+  \centering
   \begin{align*}
@@
   &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \; \middle| \;
     \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+      & n = 0 \lor (\dom(\hs) = [\start,\addrend] \land \\
+      &\forall \addr \in [\start,\addrend] \ldotp\\
+      & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
     \end{aligned}
-        \right\} \union \{\npair[0]{\ms} \}
-\end{align*}
-Revoking all temporary regions:
-\begin{align*}
-  \revokeTemp{} & : \Worlds \fun \Worlds \\
-  \revokeTemp{W} & \defeq \lambda r \ldotp 
-                   \begin{cases}
-                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
-                     W(r)                & \text{otherwise}
-                   \end{cases}
-\end{align*}
-\newcommand{\wrev}[1]{\revokeTemp{#1}}
-
-\begin{align*}
+        \right\}\\
   &\iota_{\start,\addrend}^{\nwl} : \Regions \\
@@
   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \; \middle| \;
     \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
+      &n = 0 \lor (\dom(\hs) = [\start,\addrend] \land \\
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\quad \nonlocal{\ms(\addr)} \land\\ 
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+      &\;\; \nonlocal{\ms(\addr)} \land\\ 
+      &\;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
     \end{aligned}
-        \right\} \union \{\npair[0]{\ms} \}
+        \right\}
 \end{align*}
-
+\caption{Standard regions}
+\label{fig:standard-regions}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
   \begin{align*}
-&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
+    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
     &\qquad\begin{multlined}
@@
   \end{align*}
-
   \begin{align*}
-  & \writeCond{} : (\Regions \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
   & \writeCond{}(\iota,\gl)(W) =  \\
-  & \quad \begin{aligned}[t]
-    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-    & \;\qquad W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{ and }\\
-    & \;\qquad W(r) \text{ is address-stratified } \}
-  \end{aligned}
-\end{align*}
-
-\begin{align*}
-  & \readCond{} : \Globals \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
+  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
+    \begin{aligned}
+      & \exists r \in \var{localityReg}(g,W) \ldotp \\
+      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and}}\\
+      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
+    \end{aligned} \right\}\\
   & \readCond{}(\gl)(W) =  \\
-  & \quad \begin{aligned}[t]
-    \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
-    & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-    & \;\qquad W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl \}
-  \end{aligned}
-\end{align*}
-
-\begin{align*}
+  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
+    \begin{aligned}
+      & \exists r \in \var{localityReg}(g,W) \ldotp \\
+      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+      & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
+    \end{aligned} \right\}\\
   & \execCond{}(\gl)(W) = \\
-  & \quad
-    \begin{aligned}[t]
-      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
-      & \quad \forall W' \future W \ldotp\\
-      & \qquad\forall a \in [\start,\addrend] \ldotp\\
-      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
-    \end{aligned} \\
+  & \left\{ \npair{(\var{perms},\start,\addrend)} \; \middle| \;
+    \begin{aligned}
+      & \forall n' < n, W' \future W \ldotp \\
+      & \;\; \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
+      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
+    \end{aligned} \right\} \\
+  & \entryCond{}(\gl)(W) = \\
+  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
+    \begin{aligned}
+ &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
+      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
+    \end{aligned} \right\} \\
   & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
@@
 
-\begin{align*}
-  & \entryCond{}(\gl)(W) = \\
-  & \quad
-    \begin{aligned}[t]
-      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
-      & \quad \forall W' \future W \ldotp\\
-      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
-    \end{aligned} \\
-  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
-\end{align*}
-
+\caption{Permission-based conditions}
+\label{fig:perm-conds}
+\end{figure}
 Now define the value relation as follows:

### 2017-02-08 csf-2017/paper.tex
changed land and lor to text
Minor editing of the intro.

--- 2017-02-08 csf-2017/paper.tex
@@
 Compromising software security is often based on attacks that break programming
-language properties relied on by software authors, such as control flow
+language properties relied upon by software authors, such as control flow
 integrity, local state encapsulation etc. Commodity processors offer little
@@
 scalability. As a result, defenses against attacks on control flow integrity and
-local state encapsulation necessarily rely on techniques like machine code
-rewriting (e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}) or
+local state encapsulation are either limited to only certain common forms of
+attacks (leading to an attack-defense arms race) or rely on techniques like
+machine code rewriting
+(e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
 verification (e.g.~\cite{morrisett_system_1999}), virtual machines with a native
 stack (e.g.~the JVM~\cite{lindholm_java_2014}) or
-randomization~\cite{forrest_building_1997}. Such techniques for retrofitting
-protection techniques to commodity hardware have disadvantages in terms of
-performance, complexity and/or security.
+randomization~\cite{forrest_building_1997}. The latter techniques essentially
+emulate protection techniques on existing hardware, at the cost of performance,
+system complexity and/or security.
 
 \emph{Capability machines} are a type of processors, intended to remediate these
-limitations by presenting a better security model at the hardware
-level~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740,Carter:1994:HSF:195473.195579}.\dominique{todo:
-  more?} \lars{can we add that there has been renewed interest and
-  refer to crash-safe project ?}
- Such hardware tags every word (in the register file and in memory) to
-enforce a strict separation between numbers and capabilities, a kind of pointers
-carrying some kind of authority. Memory capabilities carry the authority to read
-and/or write to a range of memory locations. Additionally, the details vary but
-all capability machines offer some form of \emph{object capabilities}, which
+limitations by presenting a better security model at the hardware level. They
+are based on old ideas
+(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
+but have recently received renewd interest with particularly the CHERI project
+proposing new ideas and tackling practical challenges like backwards
+compatibility and realistic OS
+support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
+machines tag every word (in the register file and in memory) to enforce a
+strict separation between numbers and capabilities, a kind of pointers carrying
+some kind of authority. Memory capabilities carry the authority to read and/or
+write to a range of memory locations. Additionally, the details vary but all
+capability machines offer some form of \emph{object capabilities}, which
 represent the authority to invoke a piece of code without exposing the code's

### 2017-02-08 csf-2017/paper.tex

--- 2017-02-08 csf-2017/paper.tex
@@
         &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
-        &\quad (\hs = \biguplus_{r\in\activeReg{W}}P(r) \land \\
+        &\quad (\hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\
         &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
-        &\qquad W(r) = (\_,s,\_,\_,H) \land \\
+        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \\
         &\qquad \npair[n]{P(r)} \in H(s)(\xi^{-1}(W)))\\
@@
         & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
-        & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \land \\ 
+        & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{\footnotesize{ and }} \\ 
         & \;\; \heapSat[\hs']{n-i}{W'} 
@@
       & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
       & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
@@
       & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
       & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
@@
       & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
       & \quad \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \} 
@@
       & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
       & \quad \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) 
@@
       & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
-      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
+      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
       & \quad \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \}
@@
     \begin{aligned}
-      & n = 0 \lor (\dom(\hs) = [\start,\addrend] \land \\
+      & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
       &\forall \addr \in [\start,\addrend] \ldotp\\
@@
     \begin{aligned}
-      &n = 0 \lor (\dom(\hs) = [\start,\addrend] \land \\
+      &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
       &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\;\; \nonlocal{\ms(\addr)} \land\\ 
+      &\;\; \nonlocal{\ms(\addr)} \text{\footnotesize{ and }}\\ 
       &\;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))

### 2017-02-08 csf-2017/paper.tex
Reordered figure
added figure from SCM presentation
update to LR figure

@@
 \usepackage{rotating}
+
+\usepackage{afterpage}
 
@@
   interpretation function.}
-
+\begin{figure}[htbp]
+  \centering
+  \includegraphics{w11}
+  \caption{Relation between local and global capabilities and temporary and permanent regions.}
+  \label{fig:cap-world}
+\end{figure}
 A world is a finite map from region names, modeled as natural numbers, to regions.
@@
 \subsection{Logical relation}
-
-\subsubsection{Observation relation}
-\todo[inline]{write something about single orthogonal closure?}
-
- The \emph{observation relation}, $\observations$, defines which
-configurations are capability safe. Roughly speaking,
-a configuration is capability safe if, whenever it halts, then
-the resulting memory satisfies a private future world. 
-Notice that if a configuation fails, e.g., if an unauthorized access 
-(e.g., a load of a memory location without the requisite read capability)
-is attempted, then the configuation is considered capability safe, which it should
-because then the capability machine has caught the attempted unauthorized access
-This is similar to the logical relation given for an untyped language
-in~\cite{Devriese:2016ObjCap}, but unlike logical relations for typed languages, which
-typically require that computations do not fail.
-
-\subsubsection{Expression and register-file relation}
-The expression relation, $\stder$, is a world-indexed predicate that
-contains all the words that safely can be put in the $\pcreg$-register
-of a register-file that only gives access to capabilities that are capability
-safe.
-
+\afterpage{
 \begin{figure}[htbp]
@@
 
-The register-file relation, $\stdrr$, contains all the register files
-that contains capability safe words. It does not require the word in
-the $\pcreg$-register to be capability safe, because the relation is
-only used in the expression relation, where the $\pcreg$-register is
-overwritten anyway.
-
-\subsubsection{Value relation}
-The value relation, $\stdvr$, contains all the words that are safe to
-give away to an adversary. The conditions a capability must satisfy to
-be safe to give to an adversary depends on what permissions it
-gives. We there for define four conditions
-
 \begin{figure}[htbp]
@@
   \centering
-  \begin{align*}
-    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
-    &\qquad\begin{multlined}
-      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
-      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
-      \dom(\ms) = \dom(\ms') \wedge \\
-      \forall \addr \in
-      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
-    \end{multlined}
-  \end{align*}
   \begin{align*}
@@
 \end{align*}
+  \begin{align*}
+    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
+    &\qquad\begin{multlined}
+      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
+      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
+      \dom(\ms) = \dom(\ms') \wedge \\
+      \forall \addr \in
+      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
+    \end{multlined}
+  \end{align*}
 
@@
 \end{figure}
+\clearpage
+}
+
+\subsubsection{Observation relation}
+\todo[inline]{write something about single orthogonal closure?}
+
+ The \emph{observation relation}, $\observations$, defines which
+configurations are capability safe. Roughly speaking,
+a configuration is capability safe if, whenever it halts, then
+the resulting memory satisfies a private future world. 
+Notice that if a configuation fails, e.g., if an unauthorized access 
+(e.g., a load of a memory location without the requisite read capability)
+is attempted, then the configuation is considered capability safe, which it should
+because then the capability machine has caught the attempted unauthorized access
+This is similar to the logical relation given for an untyped language
+in~\cite{Devriese:2016ObjCap}, but unlike logical relations for typed languages, which
+typically require that computations do not fail.
+
+\subsubsection{Expression and register-file relation}
+The expression relation, $\stder$, is a world-indexed predicate that
+contains all the words that safely can be put in the $\pcreg$-register
+of a register-file that only gives access to capabilities that are capability
+safe.
+
+The register-file relation, $\stdrr$, contains all the register files
+that contains capability safe words. It does not require the word in
+the $\pcreg$-register to be capability safe, because the relation is
+only used in the expression relation, where the $\pcreg$-register is
+overwritten anyway.
+
+\subsubsection{Value relation}
+The value relation, $\stdvr$, contains all the words that are safe to
+give away to an adversary. The conditions a capability must satisfy to
+be safe to give to an adversary depends on what permissions it
+gives. We there for define four conditions
+
 Now define the value relation as follows:
@@
 
-\bibliographystyle{plainnat}
-\bibliography{references}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,references}
 \end{document}

### 2017-02-08 csf-2017/paper.tex
Make a start on presenting the capability machine.

--- 2017-02-08 csf-2017/paper.tex
@@
 \newcommand{\ExecConfs}{\plaindom{ExecConf}}
-\newcommand{\RegName}{\plaindom{RegisterName}}
+\newcommand{\RegName}{\plaindom{RegName}}
 \newcommand{\Regs}{\plaindom{Reg}}
@@
 
+\newcommand{\isdef}{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}
+\newcommand\bnfdef{\mathrel{::=}}
+
 \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
@@
 \label{sec:introduction}
-\todo[inline]{See the latex comments for the next paragraphs!}
 
@@
   using a variant of an idea that has been used by
-  \cite{Dreyer:2010:IHS:1863543.1863566} to express well-bracketedness
+  \cite{Dreyer:jfp12} to express well-bracketedness
   guarantees in a lambda calculus: Kripke worlds with a weak and a strong future
@@
 
-\subsection{Introduce capability machines}
-\subsection{Enforcing properties on a capability machine}
-
-Privacy of local state using memory capabilities.
-Encapsulation of private state using enter pointers or CHERI's ccall.
-(mention in passing secure compilation, full abstraction)
-
-\subsection{Well-bracketed control flow and encapsulation of stack frames}
-
-Hard to enforce using standard capabilities:
-\begin{itemize}
-\item Well-bracketed control flow can be broken if the adversary stores return pointers.
-(alternative: trusted stack)
-\item Encapsulation of local stack frames can be broken if the adversary stores stack pointers.
-\end{itemize}
-
-\subsection{Local capabilities}
-
-\begin{itemize}
-\item Introduced by CHERI for this purpose.
-\item Explain how they can be used for passing return pointers and stack pointers.
-\item But this is not watertight.
-\item Example attack:
-\begin{itemize}
-\item In a first invocation, adversary fills the entire unused part of the stack with copies of the current stack pointer
-\item In a second invocation, adversary can load the old stack pointer of the stack
-\item He now has access to private state that was pushed on the stack in between.
-\end{itemize}
-\item So:
-\begin{itemize}
-\item Can we make it (provably) watertight?
-\item Is the trusted stack necessary? Why not just pass return pointers locally?
-\item General question: How to reason about capability machines and
-local capabilities?
-\end{itemize}
-\end{itemize}
-
-\subsection{Contributions}
-
-\begin{itemize}
-\item Formal model of a simple but representative capability machine with local capabilities
-\item Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
-\begin{itemize}
-\item untrusted adversary
-\item higher-order code: callbacks passed to and received from the adversary
-\item efficient stack management
-\item no OS support
-\end{itemize}
-\item Logical relation for reasoning about code in this capability machine
-\item Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
-\item Technical contributions: reuse existing ideas in a new way:
-\begin{itemize}
-\item replace biorthogonal closure by a single orthogonal closure
-because assembly languages remove the distinction between the
-continuation and the arguments
-\end{itemize}
-\end{itemize}
-LB: I think we should have some discussion of the flexibility /
-  strength of our LR.  The LR defines which computations we think of
-  as well-behaved.   
- We should give simple / trivial examples of
-  code not in the LR. But we may also want to emphasize that the LR is 
-  not too restricting, e.g., it does not enforce a certain calling
-  scheme (examplied by Example \verb!f1! which
-  does not use the stack and the following examples which do use the
-  stack). This ties in to the discussion of biorthogonality, see the
-  following paragraphs.
-
-LB: we need to relate this carefully to Hur-Dreyer, who used
-      biorthogonality even though they also worked with an assembly
-      language. (If I recall correctly, they assumed some properties
-      of the low level language, which is why the biortho was the
-      right ting, but we need to check.)
-
-LS: It is my impression that they were interested in the relation
-between high-level and low-level programs. They are therefore in
-particular interested in low-level programs compiled from high-level
-programs which means that the continuation is always invoked in the 
-same way. Our realisation was that anything we pass as an argument 
-in a "call" can be used as a continuation, so the continuation 
-relation was redundant. Even though we do try to give our programs
-some structure using the macros, our logical relation is strong enough
-to handle unstructured programs as well.
-\begin{itemize}
-\item STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
-\end{itemize}
-\begin{itemize}
-\item Demonstrate all of this on several examples:
-\begin{itemize}
-\item security examples
-\item the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
-\item some compartmentalisation result?
-\item whatever else we do..
-\end{itemize}
-\end{itemize}
+% \subsection{Contributions}
+
+% \begin{itemize}
+% \item Formal model of a simple but representative capability machine with local capabilities
+% \item Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
+% \begin{itemize}
+% \item untrusted adversary
+% \item higher-order code: callbacks passed to and received from the adversary
+% \item efficient stack management
+% \item no OS support
+% \end{itemize}
+% \item Logical relation for reasoning about code in this capability machine
+% \item Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
+% \item Technical contributions: reuse existing ideas in a new way:
+% \begin{itemize}
+% \item replace biorthogonal closure by a single orthogonal closure
+% because assembly languages remove the distinction between the
+% continuation and the arguments
+% \end{itemize}
+% \end{itemize}
+% LB: I think we should have some discussion of the flexibility /
+%   strength of our LR.  The LR defines which computations we think of
+%   as well-behaved.   
+%  We should give simple / trivial examples of
+%   code not in the LR. But we may also want to emphasize that the LR is 
+%   not too restricting, e.g., it does not enforce a certain calling
+%   scheme (examplied by Example \verb!f1! which
+%   does not use the stack and the following examples which do use the
+%   stack). This ties in to the discussion of biorthogonality, see the
+%   following paragraphs.
+
+% LB: we need to relate this carefully to Hur-Dreyer, who used
+%       biorthogonality even though they also worked with an assembly
+%       language. (If I recall correctly, they assumed some properties
+%       of the low level language, which is why the biortho was the
+%       right ting, but we need to check.)
+
+% LS: It is my impression that they were interested in the relation
+% between high-level and low-level programs. They are therefore in
+% particular interested in low-level programs compiled from high-level
+% programs which means that the continuation is always invoked in the 
+% same way. Our realisation was that anything we pass as an argument 
+% in a "call" can be used as a continuation, so the continuation 
+% relation was redundant. Even though we do try to give our programs
+% some structure using the macros, our logical relation is strong enough
+% to handle unstructured programs as well.
+
+% DD: I think the above comment by LS is correct: the reason why the biortho is
+% the right thing for Hur-Dreyer is that their LR only accepts well-behaved
+% programs which treat the return pointer as a continuation (i.e. don't store
+% it), while ours is more general.
+
+% \begin{itemize}
+% \item STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
+% \end{itemize}
+% \begin{itemize}
+% \item Demonstrate all of this on several examples:
+% \begin{itemize}
+% \item security examples
+% \item the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
+% \item some compartmentalisation result?
+% \item whatever else we do..
+% \end{itemize}
+% \end{itemize}
 
 \subsection{Outline}
+\todo[inline]{TODO}
 
 \section{A capability machine with local capabilities}
-
-Formalisation
+\begin{figure}
+  \begin{align*}
+    \addr   &\in& \Addrs &\isdef \nats\\
+    w &\in&\Words &\isdef \ints + \Caps \\
+    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
+    \gl&\in&\Globals & ::= \glob \mid \local \\
+     && \Caps  &::= \left\{
+                       \begin{multlined}
+                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
+                         \start\in\Addrs,\addrend \in
+                         \Addrs\cup \{\infty\}
+                       \end{multlined} \right\}\\
+    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
+    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
+    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
+    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
+    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
+    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
+  \end{align*}
+
+  \begin{array}{rcl}
+    \lv    &::=& \refreg{r} \\
+    \hv    &::=& \refheap{r}\\
+    \rv    &::=& n \mid \lv \\
+    i      &::=& 
+                 \jmp{\lv} \mid 
+                 \jnz{\lv}{\rv} \mid
+                 \move{\lv}{\rv} \mid 
+                 \load{\lv}{\hv} \mid 
+                 \store{\hv}{\rv} \mid  \\
+           &   & \plus{\lv}{\rv}{\rv} \mid 
+                 \lea{\lv}{\rv} \mid 
+                 \restricttwo{\lv}{\rv} \mid 
+                 \subseg{\lv}{\rv}{\rv} \mid  \\
+           &   & \isptr{\lv}{\rv} \mid 
+                 \getp{\lv}{\lv} \mid 
+                 \getl{\lv}{\lv} \mid 
+                 \getb{\lv}{\lv} \mid
+                 \gete{\lv}{\lv} \mid\\
+           &   & \geta{\lv}{\lv} \mid 
+                 \fail \mid
+                 \halt 
+  \end{array}
+  \caption{The syntax of our capability machine assembly language.}
+  \label{fig:syntax}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (7) {$\rwlx$};
+    \node[main node] (8) [below left of=7] {$\readwritel$};
+    \node[main node] (1) [below right of=7] {$\rwx$};
+    \node[main node] (2) [below right of=1] {$\exec$};
+    \node[main node] (3) [below right of=2] {$\entry$};
+    \node[main node] (4) [below left of=1] {$\readwrite$};
+    \node[main node] (5) [below right of=4] {$\readonly$};
+    \node[main node] (6) [below right of=5] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (7) edge (8)
+    (7) edge (1)
+    (8) edge (4)
+    (1) edge (2)
+    (2) edge (3)
+    (2) edge (5)
+    (3) edge (6)
+    (1) edge (4)
+    (4) edge (5)
+    (5) edge (6);
+  \end{tikzpicture}
+
+  \caption{Permission hierarchy}
+  \label{fig:perm-hier}
+\end{figure}
+
+We define the syntax of our capability machine in Figure~\ref{fig:syntax}. We
+assume an infinite set of addresses $\Addrs$ and define machine words as either
+integers or capabilities of the form $((\perm,\gl),\start,\addrend,\addr)$. Such
+a capability represents the authority to execute permissions $\perm$ on the
+memory array $[\start,\addrend]$, together with a current address $\addr$ and a
+tag $\gl$ which indicates whether the capability is global or local. The
+available permissions and their ordering are depicted in
+Figure~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
+readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
+read/write/execute ($\rwx$) permissions. Additionally, there are somewhat
+special read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and
+enter ($\entry$) permissions which we will explain below.
+
+We assume a set of register names $\RegName$. We define register files $\reg$ as
+functions mapping register names to words and memories as functions that map
+addresses to words. The state of the entire machine is represented as a
+configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
+and a register file, or a failed or halted state, where the latter keeps track
+of the final state of memory.
 
@@
 \section{Stack and return pointer management using local capabilities}
-
 \begin{itemize}
@@
 encapsulation about them, we define a unary logical relation.
-
 % Intuition

### 2017-02-08 csf-2017/paper.tex
text about VR, readCond, and writeCond

--- 2017-02-08 csf-2017/paper.tex
@@
   \begin{align*}
-  & \writeCond{}(\iota,\gl)(W) =  \\
-  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
-    \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and}}\\
-      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
-    \end{aligned} \right\}\\
   & \readCond{}(\gl)(W) =  \\
@@
       & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
+    \end{aligned} \right\}\\
+  & \writeCond{}(\iota,\gl)(W) =  \\
+  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
+    \begin{aligned}
+      & \exists r \in \var{localityReg}(g,W) \ldotp \\
+      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and }}\\
+      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
     \end{aligned} \right\}\\
@@
 
-\subsubsection{Observation relation}
-\todo[inline]{write something about single orthogonal closure?}
-
 The \emph{observation relation}, $\observations$, defines which
@@
 
-\subsubsection{Expression and register-file relation}
 The expression relation, $\stder$, is a world-indexed predicate that
@@
 
-\subsubsection{Value relation}
-The value relation, $\stdvr$, contains all the words that are safe to
-give away to an adversary. The conditions a capability must satisfy to
-be safe to give to an adversary depends on what permissions it
-gives. We there for define four conditions
-
-Now define the value relation as follows:
+The \emph{value relation}, $\stdvr$, defines which words are capability
+safe. A word is capability safe if it cannot be used to break the
+protocols defined by the world. Data is always capability safe because
+it provides no authority that can be used to break the
+invariants. Capabilities gives the authority to manipulate memory and
+potentially break memory protocols, so they need to satisfy certain
+conditions to be capability safe. We define a condition for each kind
+of permission a capability can have. These permission-based conditions
+can be found in \figurename~\ref{fig:perm-conds}.
+
+The condition for capabilities with read permission is
+$\readCond{}$. This condition makes sure the read capability can only
+be used to read capability safe words. To make sure this is the case,
+the region that governs the part of memory the capability grants
+authority over must at least require all the words to be in the value
+relation.
+Generally speaking, global capabilities should only depend on
+permanent regions. The function $\var{localityReg}(g,W)$ projects out
+all active regions when $g$ is local, but when $g$ is global, then it
+is just the permanent regions.
+
+For capabilities with a write permission, $\writeCond{}$ must be
+satisfied for the capabilities range of authority. Generally speaking,
+write capabilities may be used to write any word, so the region that
+governs the part of memory the write capability has authority over
+must allow any capability safe word.
+This is of course not entirely true because of the distinction between
+write and permit-write-local capabilities. For write capabilities, the
+region must require the contents of memory to be non-local and
+capability safe words. For permit-write-local capabilities, the
+non-local requirement is omitted.
+
+The write condition uses $\var{localityReg}$ to pick an appropriate
+region in the same way the read condition does.
+
+The condition for the region to be address-stratified is there for
+technical reasons.\lau{Dominique, you probably disagree with this, so
+  please feel free to add something.}
+
+The definition of the value relation in \figurename~\ref{fig:logrel}
+is defined in terms of the permission-based conditions.
 

### 2017-02-08 local-capability-machine.tex
typo

@@
 
+\newcommand{\isdef}{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}
+\newcommand\bnfdef{\mathrel{::=}}
+
 \begin{document}
@@
 \begin{align*}
-  \Addrs &::= \nats\\
-  \Words &::= \Caps + \ints \\
-  \Regs  &::= \RegName \rightarrow \Words\\
-  \Heaps &::= \Addrs \rightarrow \Words \\
-  \Perms &::= \{ \noperm, \readonly, \readwrite, \readwritel, \exec, \entry, \rwx, \rwlx\}\\
-  \ExecConfs  &::= \Regs \times \Heaps \\
-  \Globals & ::= \{\glob, \local \} \\
-  \Caps  &::= (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
-  \Confs &::= \ExecConfs + \{\failed \} + \halted \times \Heaps \\
-  \HeapSegments &::= \Addrs \parfun \Words
+  \Addrs &\isdef \nats\\
+  \Words &\isdef \Caps + \ints \\
+  \Regs  &\isdef \RegName \rightarrow \Words\\
+  \Heaps &\isdef \Addrs \rightarrow \Words \\
+  \Perms &\bnfdef  \noperm\mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
+  \ExecConfs  &\isdef \Regs \times \Heaps \\
+  \Globals & \bnfdef \glob\mid \local \\
+  \Caps  &\isdef (\Perms \times \Globals) \times \Addrs \times (\Addrs + \{ \infty \}) \times \Addrs\\
+  \Confs &\isdef \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
+  \HeapSegments &\isdef \Addrs \parfun \Words
 \end{align*}
@@
 \end{align*}
-\lau{At some point we talked about whether $\updatePcPerm{}$ should just give a $\glob$ capability all the time.
+% \lau{At some point we talked about whether $\updatePcPerm{}$ should just give a $\glob$ capability all the time.
 I believe this is dangerous as it can be used to first get a $\glob$ read-capability for the stack and subsequently in a $\glob$ callback, you can start reading from someone else's private stack!}
+% \dominique{You're right, that would be dangerous and we should not do it.}
 % TODO: \Phi.reg(rv) to some other notation. It should only look up reg, if it is a regname otherwise just the litteral.
@@
  \[
-   W \futurestr W' \Leftrightarrow \xi(W) \futurestr \xi(W')
+   W' \futurestr W \Leftrightarrow \xi(W') \futurestr \xi(W')
  \]
@@
  \[
-   W \futurewk W' \Leftrightarrow \xi(W) \futurewk \xi(W')
+   W' \futurewk W \Leftrightarrow \xi(W') \futurewk \xi(W)
  \]
@@
   \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-  \observations (W) \defeq & \lambda W \ldotp 
+  \observations \defeq & \lambda W \ldotp 
                              \{ \npair{(\reg,\hs)} \mid

### 2017-02-08 csf-2017/paper.tex
fix so document compiles
text about exec/enter cond

@@
   \end{align*}
-
+\[
   \begin{array}{rcl}
@@
   \end{array}
+\]
   \caption{The syntax of our capability machine assembly language.}
@@
 non-local requirement is omitted.
-
 The write condition uses $\var{localityReg}$ to pick an appropriate
 region in the same way the read condition does.
-
 The condition for the region to be address-stratified is there for
@@
   please feel free to add something.}
+
+The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
+differs only slightly. Executable capabilities can be updated to point
+anywhere in memory, so they must be capability safe as a program
+counter no matter where they point in memory. Contrarily, the enter
+capabilities are opaque and must be used from the address they point
+to. They change permission when jumped to, so we require them to be
+capability safe as a program counter when the permission is changed to
+read-execute.
+In both conditions, the capability must be capability safe as a
+program counter certain future worlds. If the capability is global,
+then it should be capability safe as a program counter in any private
+future world. This means that it can only depend on permanent regions
+because temporary regions may be revoked in a private future
+world. On the other hand, local capabilities must only be capability
+safe in public future worlds because they may depend on temporary
+regions, so we cannot expect them to function correctly as program
+counters if temporary regions are revoked.
 

### 2017-02-08 csf-2017/paper.tex
Added FTLR
updated the theorem to a readable version
pass over intro and logrel

@@
 language properties relied upon by software authors, such as control flow
-integrity, local state encapsulation etc. Commodity processors offer little
+integrity, local state encapsulation, etc. Commodity processors offer little
 support for defending against such attacks: they offer security primitives with
@@
 machine code rewriting
-(e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
-verification (e.g.~\cite{morrisett_system_1999}), virtual machines with a native
-stack (e.g.~the JVM~\cite{lindholm_java_2014}) or
+(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
+verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a native
+stack (e.g.,~the JVM~\cite{lindholm_java_2014}) or
 randomization~\cite{forrest_building_1997}. The latter techniques essentially
@@
 
-\emph{Capability machines} are a type of processors, intended to remediate these
-limitations by presenting a better security model at the hardware level. They
-are based on old ideas
+\emph{Capability machines} are a type of processors, intended to
+remediate these limitations by presenting a better security model at
+the hardware level. They are based on old ideas
 (e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
-but have recently received renewd interest with particularly the CHERI project
-proposing new ideas and tackling practical challenges like backwards
-compatibility and realistic OS
+but have recently received renewed interest; in particular, the CHERI
+project has proposed new ideas and ways of tackling practical
+challenges like backwards compatibility and realistic OS
 support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
-machines tag every word (in the register file and in memory) to enforce a
-strict separation between numbers and capabilities, a kind of pointers carrying
-some kind of authority. Memory capabilities carry the authority to read and/or
-write to a range of memory locations. Additionally, the details vary but all
-capability machines offer some form of \emph{object capabilities}, which
-represent the authority to invoke a piece of code without exposing the code's
-encapsulated private state (e.g. the M-Machine's enter capabilities or CHERI's
-sealed code/data pairs).
-
-Unlike for commodity processors, the security primitives on capability machines
-lend themselves well to the enforcement of properties like local state
-encapsulation and control-flow integrity. Potentially, they will enable new
-compilation schemes that enforce such properties in a way that is efficient but
-also 100\% watertight (ideally evidenced by a mathematical proof, guaranteeing
-that we do not end up in a new attack-defense arms race). However, there is
-still a lot that needs to happen to attain this goal. For example, it is quite
-non-trivial to devise a compilation scheme adapted to the details of a specific
-language's notion of encapsulation (e.g. encapsulation of private member
-variables in OO languages often behaves quite differently than encapsulation of
-private state in ML-like languages). And even if such a scheme were defined, a
-formal proof depends on a formalisation of the encapsulation provided by the
-capability machine at hand.
+machines tag every word (in the register file and in memory) to
+enforce a strict separation between numbers and capabilities, a kind
+of pointers carrying some kind of authority. Memory capabilities carry
+the authority to read and/or write to a range of memory
+locations. Moreover, all capability machines
+offer some form of \emph{object capabilities}, which represent the
+authority to invoke a piece of code without exposing the code's
+encapsulated private state (e.g., the M-Machine's enter capabilities or
+CHERI's sealed code/data pairs).
+
+In contrast to the situation for commodity processors, the security
+primitives on capability machines lend themselves well to the
+enforcement of properties like local state encapsulation and
+control-flow integrity. Potentially, they will enable new compilation
+schemes that enforce such properties in a way that is efficient but
+also 100\% watertight (ideally evidenced by a mathematical proof,
+guaranteeing that we do not end up in a new attack-defense arms
+race). However, there is still a lot that needs to happen to attain
+this goal. For example, it is quite non-trivial to devise a
+compilation scheme adapted to the details of a specific language's
+notion of encapsulation (e.g., encapsulation of private member
+variables in OO languages often behaves quite differently than
+encapsulation of private state in ML-like languages). And even if such
+a scheme were defined, a formal proof depends on a formalisation of
+the encapsulation provided by the capability machine at hand.
 
 A similar problem is the enforcement of control-flow integrity on capability
-machines. An interesting approach is taken in CheriBSD. They drop the standard
-contiguous C stack and split it into a central, trusted shadow stack, managed by
+machines. An interesting approach is taken in
+CheriBSD~\citep{Watson2015Cheri}: It drops the standard
+contiguous C stack and splits it into a central, trusted shadow stack, managed by
 trusted call and return instructions\footnote{These call and return instructions
@@
   them very fast to invoke.}, and disjoint, per-component
-stacks~\citep{Watson2015Cheri}. For preventing illegal use of stack references,
+stacks. To prevent illegal use of stack references,
 the approach relies on \emph{local capabilities}, a type of capabilities offered
@@
 the duration of the invocation, and revoked afterwards. However, a lot of
-details remain to be filled in (e.g. how to combine a trusted shadow stack with
+details remain to be investigated (e.g., how to combine a trusted shadow stack with
 (potentially) cross-domain function pointers in a scalable way?) and there is no
-argument that this enforcement of CFI is watertight.
-
-In this paper, we make progress towards the goal of efficient, provably
+argument that this enforcement of control-flow integrity is watertight.
+
+In this paper, we make progress towards the goal of efficient and provably
 watertight enforcement of encapsulation and control-flow integrity on capability
@@
   call and return instructions. It also does not require separate per-component
-  stacks and instead works with a single, fairly standard contiguous stack. It
-  supports higher-order cross-component calls (e.g. (potentially)
-  cross-component function pointers) and can be efficient assuming one
-  additional piece of processor support: an instruction for efficiently clearing
+  stacks and instead works with a single, fairly standard, contiguous stack. It
+  supports higher-order cross-component calls (e.g., (potentially)
+  cross-component function pointers) and can be efficient assuming
+  only \lars{can we say only here ?}
+  one additional piece of processor support: an instruction for efficiently clearing
   a range of memory.
-\item We provide a theorem (capability-safety) that is (by far) the most general
+\item We provide a theorem that is (by far) the most general
   and powerful formulation of the guarantees offered by a capability machine.
@@
   specific way of using the system's capabilities.
-\item The above theorem is formulated in terms of a unary step-indexed logical
-  relation with recursively defined Kripke worlds. It is inspired by one used by
-  \cite{Devriese:2016ObjCap} for capability-safety in a JavaScript-like lambda
-  calculus, but has two novel technical characteristics. The natural CPS style
-  of assembly code with return pointers as continuations is modeled using a
-  \emph{single} orthogonal closure, rather than the biorthogonal closure that
-  has been previously used for lambda calculi. Perhaps more surprisingly, we
-  show that the special nature of local capabilities can be elegantly expressed
-  using a variant of an idea that has been used by
-  \cite{Dreyer:jfp12} to express well-bracketedness
-  guarantees in a lambda calculus: Kripke worlds with a weak and a strong future
-  world relations. Interestingly, the two future world relations generally work
-  very similarly to Dreyer et al.'s, except that they turn up in quite a
-  different place in the logical relation.
+\item The above theorem is formulated in terms of a unary step-indexed
+  logical relation with recursively defined Kripke worlds. It is
+  inspired by one used in \cite{Devriese:2016ObjCap} for
+  capability-safety in a JavaScript-like lambda calculus, but has two
+  novel technical characteristics. The natural CPS style of assembly
+  code with return pointers as continuations is modeled using a
+  \emph{single} orthogonal closure, rather than the biorthogonal
+  closure that has been previously used for lambda calculi. Perhaps
+  more surprisingly, we show that the special nature of local
+  capabilities can be elegantly expressed using a variant of an idea
+  that has been used by Dreyer et. al.~\cite{Dreyer:jfp12} to express
+  well-bracketedness guarantees in a lambda calculus: Kripke worlds
+  with both a weak and a strong future world relation. Interestingly,
+  the two future world relations generally work very similarly to
+  Dreyer et al.'s, except that they turn up in quite a different place
+  in the logical relation.
 \item We demonstrate the power of our results by applying them to challenging
@@
 \end{itemize}
+
+For reasons of space, som details and all proofs have been omitted;
+please refer to the technical appendix~\cite{} for those.
 
@@
 The condition for capabilities with read permission is
-$\readCond{}$. This condition makes sure the read capability can only
-be used to read capability safe words. To make sure this is the case,
-the region that governs the part of memory the capability grants
-authority over must at least require all the words to be in the value
-relation.
-Generally speaking, global capabilities should only depend on
-permanent regions. The function $\var{localityReg}(g,W)$ projects out
-all active regions when $g$ is local, but when $g$ is global, then it
-is just the permanent regions.
-
-For capabilities with a write permission, $\writeCond{}$ must be
-satisfied for the capabilities range of authority. Generally speaking,
+$\readCond{}$. It ensures that a read capability can only be used to
+read capability safe words, by requiring that the region that governs
+the part of memory that the capability grants authority over enforces
+that all the words are in the value relation.  Generally speaking,
+global capabilities should only depend on permanent regions. Therefore
+we use the function $\var{localityReg}(g,W)$, which projects out all
+active regions when $g$ is local, but only the permanent regions when
+$g$ is global.
+
+For a capability with a write permission, $\writeCond{}$ must be
+satisfied for the capability's range of authority. Generally speaking,
 write capabilities may be used to write any word, so the region that
 governs the part of memory the write capability has authority over
-must allow any capability safe word.
+must allow any capability safe word to be stored.
 This is of course not entirely true because of the distinction between
 write and permit-write-local capabilities. For write capabilities, the
-region must require the contents of memory to be non-local and
-capability safe words. For permit-write-local capabilities, the
+region must require the contents of memory to be non-local (and
+capability safe). For permit-write-local capabilities, the
 non-local requirement is omitted.
@@
 The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
-differs only slightly. Executable capabilities can be updated to point
+differ only slightly. Executable capabilities can be updated to point
 anywhere in memory, so they must be capability safe as a program
-counter no matter where they point in memory. Contrarily, the enter
+counter no matter where they point in memory. In contrast, enter
+\lars{``capability safe as a program counter'' is a bit hard to
+  understand - why not say that this is a capability one can jump to ?}
 capabilities are opaque and must be used from the address they point
@@
 In both conditions, the capability must be capability safe as a
-program counter certain future worlds. If the capability is global,
+program counter in certain future worlds. If the capability is global,
 then it should be capability safe as a program counter in any private
@@
 \subsection{Fundamental theorem of logical relations}
+\begin{theorem}
+  \label{thm:ftlr}
+  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
+  If one of the following holds:
+  \begin{align*}
+      & \bullet
+        \begin{aligned}[t]
+        &\perm = \exec \text{\footnotesize{ and }}\\
+        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+      \end{aligned} \\
+    & \bullet 
+      \begin{aligned}[t]
+        &\perm = \rwx \text{\footnotesize{ and }} \\
+        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
+        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
+      \end{aligned} \\
+    & \bullet 
+      \begin{aligned}[t]
+        &\perm = \rwlx \text{\footnotesize{ and }}\\
+        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
+        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
+      \end{aligned}
+  \end{align*}
+  then
+  \[
+    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
+  \]
+\end{theorem}
 

### 2017-02-08 csf-2017/paper.tex
update to expression relation description

@@
 
-The expression relation, $\stder$, is a world-indexed predicate that
-contains all the words that safely can be put in the $\pcreg$-register
-of a register-file that only gives access to capabilities that are capability
-safe.
-
-The register-file relation, $\stdrr$, contains all the register files
-that contains capability safe words. It does not require the word in
-the $\pcreg$-register to be capability safe, because the relation is
-only used in the expression relation, where the $\pcreg$-register is
-overwritten anyway.
+The \emph{register-file relation}, $\stdrr$, specifies all the register-files that are capability safe.
+Capability safe register-files contain capability safe words (e.g.n, words in $\stdvr$) in all registers but $\pcreg$.
+
+The \emph{expression relation}, $\stder$, defines what words are capability safe when used for the program counter.
+A word is capability safe as a program counter when it can be plugged into a capability safe register file (i.e., a register file in $\stdrr$) and paired with a memory satisfying the world to become a capability safe configuration.
 
@@
   understand - why not say that this is a capability one can jump to ?}
+\lau{ When jumping to a capability, it becomes the program counter.
+A program counter is capability safe when it is the $\stder$ relation.
+When I say ``capability safe as a program counter'' I take it to mean being in the $\stder$-relation.
+}
 capabilities are opaque and must be used from the address they point
@@
 \subsection{Fundamental theorem of logical relations}
+
 \begin{theorem}
   \label{thm:ftlr}
-  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$  \\
-  If one of the following holds:
+  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
   \begin{align*}

### 2017-02-08 csf-2017/paper.tex
small corrections in the LR part

@@
 governs the part of memory the write capability has authority over
-must allow any capability safe word to be stored.
+must at least allow any capability safe word to be stored.
 This is of course not entirely true because of the distinction between
@@
 
-\begin{theorem}
+\begin{theorem}[Fundamental Theorem]
   \label{thm:ftlr}

### 2017-02-08 csf-2017/paper.tex
changed weak/strong to public/private in intro

@@
   well-bracketedness guarantees in a lambda calculus: Kripke worlds
-  with both a weak and a strong future world relation. Interestingly,
+  with both a public and a private future world relation. Interestingly,
   the two future world relations generally work very similarly to

### 2017-02-08 csf-2017/paper.tex
a bit of text about the FTLR

@@
 
-\subsection{Fundamental theorem of logical relations}
+With the logical relation defined, we can now state a strong theorem
+about the safety properties of the capability machine. Namely that any
+capability that does not grant unreasonable authority is capability
+safe as a program counter.
 
@@
 \end{theorem}
+When proving correctness of properties of programs, the Fundamental
+Theorem serves as a way to reason about the execution of unknown
+programs. For instance if the capability for a program only grants access to
+instructions and data, then the capability is capability safe as a
+program counter. The theorem also serves as a sanity check of the
+logical relation as it makes sure that a large class of capabilities
+actually inhabits it.
 

### 2017-02-08 csf-2017/paper.tex
update to FTLR text

--- 2017-02-08 csf-2017/paper.tex
@@
 \end{theorem}
+The Fundamental Theorem allows us to prove capability safety for
+untrusted programs. If we can show that the authority the capability
+for the program grants is reasonable, then it is safe to
+execute. Intuitively this is because the logical relation captures the
+safe behaviour of the capability machine that do not allow you to
+produce something capability \emph{unsafe} from something capability
+safe. The Funamental Theorem also serves as a sanity check of the
+logical relation as it makes sure that a large class of capabilities
+actually inhabits it.
+
+% Comparisson of LR to other work:
+% - Single othorgonal closure
+%   o all capabilities are possible continuations, not just the one we
+%   intend to be it.
+% - No termination guarantees
 

### 2017-02-08 csf-2017/paper.tex
spelling

@@
   novel technical characteristics. The natural CPS style of assembly
-  code with return pointers as continuations is modeled using a
+  code with return pointers as continuations is modelled using a
   \emph{single} orthogonal closure, rather than the biorthogonal
@@
 
-For reasons of space, som details and all proofs have been omitted;
+For reasons of space, some details and all proofs have been omitted;
 please refer to the technical appendix~\cite{} for those.
@@
 \end{figure}
-A world is a finite map from region names, modeled as natural numbers, to regions.
+A world is a finite map from region names, modelled as natural numbers, to regions.
 We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
@@
 revoked. This means that local capabilities safely can depend on the
-protocols imposed on memory by temporary regions, wheareas global
+protocols imposed on memory by temporary regions, whereas global
 capabilities cannot, since a global capability may stay around
@@
 As explained above, this can be done using a special revoked region,
-which does not desribe any memory, but we also allow  
+which does not describe any memory, but we also allow  
 any other region to serve as a mask, and thus any region $r$ can be a
@@
 }
+\todo[inline]{Add section introduction}
 
@@
 the resulting memory satisfies a private future world. 
-Notice that if a configuation fails, e.g., if an unauthorized access 
+Notice that if a configuration fails, e.g., if an unauthorized access 
 (e.g., a load of a memory location without the requisite read capability)
-is attempted, then the configuation is considered capability safe, which it should
+is attempted, then the configuration is considered capability safe, which it should
 because then the capability machine has caught the attempted unauthorized access
@@
 
-The \emph{register-file relation}, $\stdrr$, specifies all the register-files that are capability safe.
-Capability safe register-files contain capability safe words (e.g.n, words in $\stdvr$) in all registers but $\pcreg$.
-
-The \emph{expression relation}, $\stder$, defines what words are capability safe when used for the program counter.
-A word is capability safe as a program counter when it can be plugged into a capability safe register file (i.e., a register file in $\stdrr$) and paired with a memory satisfying the world to become a capability safe configuration.
+The \emph{register-file relation}, $\stdrr$, specifies all the
+register-files that are capability safe. Capability safe
+register-files contain capability safe words (e.g.n, words in
+$\stdvr$) in all registers but $\pcreg$.
+
+The \emph{expression relation}, $\stder$, defines what words are
+capability safe when used for the program counter. A word is
+capability safe as a program counter when it can be plugged into a
+capability safe register file (i.e., a register file in $\stdrr$) and
+paired with a memory satisfying the world to become a capability safe
+configuration.
 
@@
 safe as a program counter.
-
 \begin{theorem}[Fundamental Theorem]
@@
 produce something capability \emph{unsafe} from something capability
-safe. The Funamental Theorem also serves as a sanity check of the
+safe. The Fundamental Theorem also serves as a sanity check of the
 logical relation as it makes sure that a large class of capabilities
@@
 
+\todo[inline]{Add discussion/comparison to other LRs}
 % Comparisson of LR to other work:

### 2017-02-09 csf-2017/paper.tex
comments for LR discusion (to be written)
Fix compilation.

--- 2017-02-08 csf-2017/paper.tex
@@
 language properties relied upon by software authors, such as control flow
-integrity, local state encapsulation, etc. Commodity processors offer little
+integrity, local state encapsulation etc. Commodity processors offer little
 support for defending against such attacks: they offer security primitives with
@@
 machine code rewriting
-(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
-verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a native
-stack (e.g.,~the JVM~\cite{lindholm_java_2014}) or
+(e.g.~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
+verification (e.g.~\cite{morrisett_system_1999}), virtual machines with a native
+stack (e.g.~the JVM~\cite{lindholm_java_2014}) or
 randomization~\cite{forrest_building_1997}. The latter techniques essentially
@@
 
-\emph{Capability machines} are a type of processors, intended to
-remediate these limitations by presenting a better security model at
-the hardware level. They are based on old ideas
+\emph{Capability machines} are a type of processors, intended to remediate these
+limitations by presenting a better security model at the hardware level. They
+are based on old ideas
 (e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
-but have recently received renewed interest; in particular, the CHERI
-project has proposed new ideas and ways of tackling practical
-challenges like backwards compatibility and realistic OS
+but have recently received renewd interest with particularly the CHERI project
+proposing new ideas and tackling practical challenges like backwards
+compatibility and realistic OS
 support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
-machines tag every word (in the register file and in memory) to
-enforce a strict separation between numbers and capabilities, a kind
-of pointers carrying some kind of authority. Memory capabilities carry
-the authority to read and/or write to a range of memory
-locations. Moreover, all capability machines
-offer some form of \emph{object capabilities}, which represent the
-authority to invoke a piece of code without exposing the code's
-encapsulated private state (e.g., the M-Machine's enter capabilities or
-CHERI's sealed code/data pairs).
-
-In contrast to the situation for commodity processors, the security
-primitives on capability machines lend themselves well to the
-enforcement of properties like local state encapsulation and
-control-flow integrity. Potentially, they will enable new compilation
-schemes that enforce such properties in a way that is efficient but
-also 100\% watertight (ideally evidenced by a mathematical proof,
-guaranteeing that we do not end up in a new attack-defense arms
-race). However, there is still a lot that needs to happen to attain
-this goal. For example, it is quite non-trivial to devise a
-compilation scheme adapted to the details of a specific language's
-notion of encapsulation (e.g., encapsulation of private member
-variables in OO languages often behaves quite differently than
-encapsulation of private state in ML-like languages). And even if such
-a scheme were defined, a formal proof depends on a formalisation of
-the encapsulation provided by the capability machine at hand.
+machines tag every word (in the register file and in memory) to enforce a
+strict separation between numbers and capabilities, a kind of pointers carrying
+some kind of authority. Memory capabilities carry the authority to read and/or
+write to a range of memory locations. Additionally, the details vary but all
+capability machines offer some form of \emph{object capabilities}, which
+represent the authority to invoke a piece of code without exposing the code's
+encapsulated private state (e.g. the M-Machine's enter capabilities or CHERI's
+sealed code/data pairs).
+
+Unlike for commodity processors, the security primitives on capability machines
+lend themselves well to the enforcement of properties like local state
+encapsulation and control-flow integrity. Potentially, they will enable new
+compilation schemes that enforce such properties in a way that is efficient but
+also 100\% watertight (ideally evidenced by a mathematical proof, guaranteeing
+that we do not end up in a new attack-defense arms race). However, there is
+still a lot that needs to happen to attain this goal. For example, it is quite
+non-trivial to devise a compilation scheme adapted to the details of a specific
+language's notion of encapsulation (e.g. encapsulation of private member
+variables in OO languages often behaves quite differently than encapsulation of
+private state in ML-like languages). And even if such a scheme were defined, a
+formal proof depends on a formalisation of the encapsulation provided by the
+capability machine at hand.
 
 A similar problem is the enforcement of control-flow integrity on capability
-machines. An interesting approach is taken in
-CheriBSD~\citep{Watson2015Cheri}: It drops the standard
-contiguous C stack and splits it into a central, trusted shadow stack, managed by
+machines. An interesting approach is taken in CheriBSD. They drop the standard
+contiguous C stack and split it into a central, trusted shadow stack, managed by
 trusted call and return instructions\footnote{These call and return instructions
@@
   them very fast to invoke.}, and disjoint, per-component
-stacks. To prevent illegal use of stack references,
+stacks~\citep{Watson2015Cheri}. For preventing illegal use of stack references,
 the approach relies on \emph{local capabilities}, a type of capabilities offered
@@
 the duration of the invocation, and revoked afterwards. However, a lot of
-details remain to be investigated (e.g., how to combine a trusted shadow stack with
+details remain to be filled in (e.g. how to combine a trusted shadow stack with
 (potentially) cross-domain function pointers in a scalable way?) and there is no
-argument that this enforcement of control-flow integrity is watertight.
-
-In this paper, we make progress towards the goal of efficient and provably
+argument that this enforcement of CFI is watertight.
+
+In this paper, we make progress towards the goal of efficient, provably
 watertight enforcement of encapsulation and control-flow integrity on capability
@@
   call and return instructions. It also does not require separate per-component
-  stacks and instead works with a single, fairly standard, contiguous stack. It
-  supports higher-order cross-component calls (e.g., (potentially)
-  cross-component function pointers) and can be efficient assuming
-  only \lars{can we say only here ?}
-  one additional piece of processor support: an instruction for efficiently clearing
+  stacks and instead works with a single, fairly standard contiguous stack. It
+  supports higher-order cross-component calls (e.g. (potentially)
+  cross-component function pointers) and can be efficient assuming one
+  additional piece of processor support: an instruction for efficiently clearing
   a range of memory.
-\item We provide a theorem that is (by far) the most general
+\item We provide a theorem (capability-safety) that is (by far) the most general
   and powerful formulation of the guarantees offered by a capability machine.
@@
   specific way of using the system's capabilities.
-\item The above theorem is formulated in terms of a unary step-indexed
-  logical relation with recursively defined Kripke worlds. It is
-  inspired by one used in \cite{Devriese:2016ObjCap} for
-  capability-safety in a JavaScript-like lambda calculus, but has two
-  novel technical characteristics. The natural CPS style of assembly
-  code with return pointers as continuations is modeled using a
-  \emph{single} orthogonal closure, rather than the biorthogonal
-  closure that has been previously used for lambda calculi. Perhaps
-  more surprisingly, we show that the special nature of local
-  capabilities can be elegantly expressed using a variant of an idea
-  that has been used by Dreyer et. al.~\cite{Dreyer:jfp12} to express
-  well-bracketedness guarantees in a lambda calculus: Kripke worlds
-  with both a public and a private future world relation. Interestingly,
-  the two future world relations generally work very similarly to
-  Dreyer et al.'s, except that they turn up in quite a different place
-  in the logical relation.
+\item The above theorem is formulated in terms of a unary step-indexed logical
+  relation with recursively defined Kripke worlds. It is inspired by one used by
+  \cite{Devriese:2016ObjCap} for capability-safety in a JavaScript-like lambda
+  calculus, but has two novel technical characteristics. The natural CPS style
+  of assembly code with return pointers as continuations is modeled using a
+  \emph{single} orthogonal closure, rather than the biorthogonal closure that
+  has been previously used for lambda calculi. Perhaps more surprisingly, we
+  show that the special nature of local capabilities can be elegantly expressed
+  using a variant of an idea that has been used by
+  \cite{Dreyer:jfp12} to express well-bracketedness
+  guarantees in a lambda calculus: Kripke worlds with a weak and a strong future
+  world relations. Interestingly, the two future world relations generally work
+  very similarly to Dreyer et al.'s, except that they turn up in quite a
+  different place in the logical relation.
 \item We demonstrate the power of our results by applying them to challenging
@@
 \end{itemize}
-
-For reasons of space, som details and all proofs have been omitted;
-please refer to the technical appendix~\cite{} for those.
 
@@
   \end{align*}
-\[
+
+  \begin{equation*}
   \begin{array}{rcl}
@@
   \end{array}
-\]
+\end{equation*}
   \caption{The syntax of our capability machine assembly language.}
@@
   \begin{align*}
+  & \writeCond{}(\iota,\gl)(W) =  \\
+  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
+    \begin{aligned}
+      & \exists r \in \var{localityReg}(g,W) \ldotp \\
+      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
+      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and}}\\
+      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
+    \end{aligned} \right\}\\
   & \readCond{}(\gl)(W) =  \\
@@
       & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
-    \end{aligned} \right\}\\
-  & \writeCond{}(\iota,\gl)(W) =  \\
-  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
-    \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and }}\\
-      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
     \end{aligned} \right\}\\
@@
 
+\subsubsection{Observation relation}
+\todo[inline]{write something about single orthogonal closure?}
+
  The \emph{observation relation}, $\observations$, defines which
@@
 
-The \emph{register-file relation}, $\stdrr$, specifies all the register-files that are capability safe.
-Capability safe register-files contain capability safe words (e.g.n, words in $\stdvr$) in all registers but $\pcreg$.
-
-The \emph{expression relation}, $\stder$, defines what words are capability safe when used for the program counter.
-A word is capability safe as a program counter when it can be plugged into a capability safe register file (i.e., a register file in $\stdrr$) and paired with a memory satisfying the world to become a capability safe configuration.
-
-The \emph{value relation}, $\stdvr$, defines which words are capability
-safe. A word is capability safe if it cannot be used to break the
-protocols defined by the world. Data is always capability safe because
-it provides no authority that can be used to break the
-invariants. Capabilities gives the authority to manipulate memory and
-potentially break memory protocols, so they need to satisfy certain
-conditions to be capability safe. We define a condition for each kind
-of permission a capability can have. These permission-based conditions
-can be found in \figurename~\ref{fig:perm-conds}.
-
-The condition for capabilities with read permission is
-$\readCond{}$. It ensures that a read capability can only be used to
-read capability safe words, by requiring that the region that governs
-the part of memory that the capability grants authority over enforces
-that all the words are in the value relation.  Generally speaking,
-global capabilities should only depend on permanent regions. Therefore
-we use the function $\var{localityReg}(g,W)$, which projects out all
-active regions when $g$ is local, but only the permanent regions when
-$g$ is global.
-
-For a capability with a write permission, $\writeCond{}$ must be
-satisfied for the capability's range of authority. Generally speaking,
-write capabilities may be used to write any word, so the region that
-governs the part of memory the write capability has authority over
-must at least allow any capability safe word to be stored.
-This is of course not entirely true because of the distinction between
-write and permit-write-local capabilities. For write capabilities, the
-region must require the contents of memory to be non-local (and
-capability safe). For permit-write-local capabilities, the
-non-local requirement is omitted.
-The write condition uses $\var{localityReg}$ to pick an appropriate
-region in the same way the read condition does.
-The condition for the region to be address-stratified is there for
-technical reasons.\lau{Dominique, you probably disagree with this, so
-  please feel free to add something.}
-
-The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
-differ only slightly. Executable capabilities can be updated to point
-anywhere in memory, so they must be capability safe as a program
-counter no matter where they point in memory. In contrast, enter
-\lars{``capability safe as a program counter'' is a bit hard to
-  understand - why not say that this is a capability one can jump to ?}
-\lau{ When jumping to a capability, it becomes the program counter.
-A program counter is capability safe when it is the $\stder$ relation.
-When I say ``capability safe as a program counter'' I take it to mean being in the $\stder$-relation.
-}
-capabilities are opaque and must be used from the address they point
-to. They change permission when jumped to, so we require them to be
-capability safe as a program counter when the permission is changed to
-read-execute.
-In both conditions, the capability must be capability safe as a
-program counter in certain future worlds. If the capability is global,
-then it should be capability safe as a program counter in any private
-future world. This means that it can only depend on permanent regions
-because temporary regions may be revoked in a private future
-world. On the other hand, local capabilities must only be capability
-safe in public future worlds because they may depend on temporary
-regions, so we cannot expect them to function correctly as program
-counters if temporary regions are revoked.
-
-The definition of the value relation in \figurename~\ref{fig:logrel}
-is defined in terms of the permission-based conditions.
-
-With the logical relation defined, we can now state a strong theorem
-about the safety properties of the capability machine. Namely that any
-capability that does not grant unreasonable authority is capability
-safe as a program counter.
-
-\begin{theorem}[Fundamental Theorem]
-  \label{thm:ftlr}
-  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
-  \begin{align*}
-      & \bullet
-        \begin{aligned}[t]
-        &\perm = \exec \text{\footnotesize{ and }}\\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-      \end{aligned} \\
-    & \bullet 
-      \begin{aligned}[t]
-        &\perm = \rwx \text{\footnotesize{ and }} \\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
-        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
-      \end{aligned} \\
-    & \bullet 
-      \begin{aligned}[t]
-        &\perm = \rwlx \text{\footnotesize{ and }}\\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
-        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
-      \end{aligned}
-  \end{align*}
-  then
-  \[
-    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
-  \]
-\end{theorem}
-When proving correctness of properties of programs, the Fundamental
-Theorem serves as a way to reason about the execution of unknown
-programs. For instance if the capability for a program only grants access to
-instructions and data, then the capability is capability safe as a
-program counter. The theorem also serves as a sanity check of the
-logical relation as it makes sure that a large class of capabilities
-actually inhabits it.
+\subsubsection{Expression and register-file relation}
+The expression relation, $\stder$, is a world-indexed predicate that
+contains all the words that safely can be put in the $\pcreg$-register
+of a register-file that only gives access to capabilities that are capability
+safe.
+
+The register-file relation, $\stdrr$, contains all the register files
+that contains capability safe words. It does not require the word in
+the $\pcreg$-register to be capability safe, because the relation is
+only used in the expression relation, where the $\pcreg$-register is
+overwritten anyway.
+
+\subsubsection{Value relation}
+The value relation, $\stdvr$, contains all the words that are safe to
+give away to an adversary. The conditions a capability must satisfy to
+be safe to give to an adversary depends on what permissions it
+gives. We there for define four conditions
+
+Now define the value relation as follows:
+
+\subsection{Fundamental theorem of logical relations}
 

### 2017-02-09 csf-2017/paper.tex
paper update

--- 2017-02-08 csf-2017/paper.tex
@@
 
-We define the syntax of our capability machine in Figure~\ref{fig:syntax}. We
+We define the syntax of our capability machine in \figurename~\ref{fig:syntax}. We
 assume an infinite set of addresses $\Addrs$ and define machine words as either
@@
 available permissions and their ordering are depicted in
-Figure~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
+\figurename~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
 readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
@@
 revoked.
+\todo[inline]{Insert reference to figure.}
 
@@
 capability safe as a program counter when the permission is changed to
-read-execute.
+read/execute.
 In both conditions, the capability must be capability safe as a
@@
 %   intend to be it.
-% - No termination guarantees
+% - No termination guarantees in observation relation
+% - Private/public transitions used like in Dreyer-Neis-Birkedal, but
+%   also to  handle the difference in what local and global
+%   capabilities can handle.
 

### 2017-02-09 csf-2017/paper.tex
LR subsection intro added

@@
 }
-\todo[inline]{Add section introduction}
+%subsection introduction
+The logical relation defines semantically what it means to be
+capability safe. In large parts, capability safety is defined by
+when a capability safely can be given to an untrusted piece of
+code. In the following paragraphs, we will go through the definitions
+of the logical relation and provide an intuition why they define
+capability safety.
+The definition of the logical relation is found in
+\figurename~\ref{fig:logrel}, \figurename~\ref{fig:standard-regions},
+and \figurename~\ref{fig:perm-conds}.
 

### 2017-02-09 csf-2017/paper.tex
added a sentence to exec/enter cond description

@@
 $g$ is global.
+\todo[inline]{Do we want to mention the standard regions here? (They
+  are not mentioned anywhere)}
 
@@
 read/execute.
-In both conditions, the capability must be capability safe as a
+In both conditions, it must be safe to jump to the capability. This
+means that the capability must be capability safe as a
 program counter in certain future worlds. If the capability is global,

### 2017-02-09 csf-2017/paper.tex
added a bit more to LR introduction
Some work on the capability machine section.

--- 2017-02-09 csf-2017/paper.tex
@@
 \usepackage{hyperref}
+\usepackage{multicol}
 \usepackage{natbib}
@@
 \section{A capability machine with local capabilities}
+In this paper, we work with a formal capability machine that has all the
+characteristics of real capability machines, as well as local capabilities much
+like CHERI's, but is otherwise kept as simple as possible. It is inspired by
+both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
+CHERI~\cite{Watson2015Cheri}. To avoid getting bogged down in uninteresting
+detail, we assume an infinite address space and unbounded integers.
+
 \begin{figure}
@@
   \end{align*}
-\[
+  \begin{equation*}
   \begin{array}{rcl}
@@
 
-We define the syntax of our capability machine in \figurename~\ref{fig:syntax}. We
+\begin{figure*}
+  \centering
+  \begin{equation*}
+    \Phi  \rightarrow
+    \begin{cases}
+      \sem{\decode(\memheap(\addr))}(\Phi) & \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ } \\
+      \failed                                 & \text{otherwise}
+    \end{cases}
+  \end{equation*}%
+  \begin{align*}
+    \updatePcPerm{w} &=
+                       \begin{cases}
+                         ((\exec,\gl),\start,\addrend,\addr) & \text{if $w = ((\entry,\gl),\start,\addrend,\addr)$}\\
+                         w & \text{otherwise}
+                       \end{cases}\\
+    \stdUpdatePc{\Phi} &=
+                         \begin{cases}
+                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
+                           \begin{array}[t]{l}
+                             \text{if $\memreg(\pcreg) = \stdcap$}\\
+                             \quad\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}\\
+                           \end{array}
+                         \end{cases}
+  \end{align*}
+
+  \begin{tabular}{|c|p{3.7cm}|p{10.5cm}|}
+    \hline
+    $i$&$\sem{i}(\Phi)$&Conditions\\
+    \hline
+    $\fail$&$\failed$&\\
+    \hline
+    $\halt$&$(\halted,\memheap)$&\\
+    \hline
+    $\move{\refreg{r_1}}{\rv}$& $\begin{multlined}
+      \stdUpdatePc{}(\updateReg{r_1}{\\
+          \memreg(\rv)}) \end{multlined}$&\\
+    \hline
+    $\jmp{\lv}$&$\begin{multlined}
+                   \updateReg{\pcreg}{\\\updatePcPerm{\memreg(\lv)}}
+                 \end{multlined}$&\\
+    \hline
+    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$  and  $\start \leq \addr \leq \addrend$\\
+    \hline
+    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and  $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
+                                                                                       and if  $\var{w} = ((\_,\local),\_,\_,\_)$, then  $\perm \in \{\rwlx,\readwritel \}$\\
+    \hline
+    $\restricttwo{\refreg{r}}{\rv}$&$\begin{multlined}
+                                       \stdUpdatePc{}(\updateReg{r}{\\(\permp',\start,\addrend,\addr)})
+                                     \end{multlined}$  &
+                                                      $\memreg(r) = \stdcap[\permp]$ and $\permp' = \decodePermPair{\memreg(\rv)}$
+                                                      and $\permp'\sqsubseteq \permp$\\
+    \hline
+    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
+                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
+    \hline
+    \multicolumn{3}{|c|}{$\cdots$}\\
+    \hline
+    \_&$\failed$&otherwise\\
+    \hline
+  \end{tabular}
+  \caption{An excerpt from the operational semantics}
+  \label{fig:op-sem}
+\end{figure*}
+
+We define the syntax of our capability machine in Figure~\ref{fig:syntax}. We
 assume an infinite set of addresses $\Addrs$ and define machine words as either
@@
 available permissions and their ordering are depicted in
-\figurename~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
+Figure~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
 readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
-read/write/execute ($\rwx$) permissions. Additionally, there are somewhat
-special read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and
-enter ($\entry$) permissions which we will explain below.
+read/write/execute ($\rwx$) permissions. Additionally, there are three special
+read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter
+($\entry$) permissions which we will explain below.
 
@@
 configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
-and a register file, or a failed or halted state, where the latter keeps track
-of the final state of memory.
+and a register file, or a failed or halted state, where the latter keeps hold
+of the final state of memory. We also define memory segments $\ms$:
+parts of memory represented as partial functions mapping addresses to words.
+
+The instruction set of our machine is rather basic.  Instructions $i$ include
+jumps ($\jmp{}$), conditional jumps ($\jnz{}{}$), a move instruction that copies
+a word from one register to another ($\move{}{}$), load and store instructions
+for reading from and writing to memory ($\load{}{}$ and $\store{}{}$)
 
@@
 revoked.
-\todo[inline]{Insert reference to figure.}
 
@@
 capability safe as a program counter when the permission is changed to
-read/execute.
+read-execute.
 In both conditions, the capability must be capability safe as a

### 2017-02-09 csf-2017/paper.tex

@@
 }
-\todo[inline]{Add section introduction}
+%subsection introduction
+The logical relation defines semantically what it means to be
+capability safe. In large parts, capability safety is defined by
+when a capability safely can be given to an untrusted piece of
+code. In the following paragraphs, we will go through the definitions
+of the logical relation and provide an intuition why they define
+capability safety.
+The definition of the logical relation is found in
+\figurename~\ref{fig:logrel}, \figurename~\ref{fig:standard-regions},
+and \figurename~\ref{fig:perm-conds}. Some definitions are omitted and
+explained when needed. The precise definitions can be found in the
+technical report.\todo[inline]{add reference to technical report}
 
@@
 capability safe as a program counter when the permission is changed to
-read-execute.
-In both conditions, the capability must be capability safe as a
+read/execute.
+In both conditions, it must be safe to jump to the capability. This
+means that the capability must be capability safe as a
 program counter in certain future worlds. If the capability is global,

### 2017-02-09 csf-2017/paper.tex
Added four program examples

--- 2017-02-09 csf-2017/paper.tex
@@
 throughout the execution of a program, but a temporary region may be
-revoked.
-\todo[inline]{Insert reference to figure.}
+revoked. This is illustrated in \figurename~\ref{fig:cap-world}.
 
@@
 \figurename~\ref{fig:logrel}, \figurename~\ref{fig:standard-regions},
-and \figurename~\ref{fig:perm-conds}.
+and \figurename~\ref{fig:perm-conds}. Some definitions are omitted and
+explained when needed. The precise definitions can be found in the
+technical report.\todo[inline]{add reference to technical report}
 
@@
 \section{Examples}
+\begin{figure}[htbp]
+  \centering
+\begin{lstlisting}
+f1: malloc $r_l$ 1
+    store $r_l$ 1
+    fetch $r_1$ $\adv$
+    call $r_1$($[]$,$[r_l]$)
+    assert $r_l$ 1
+    halt
+\end{lstlisting}
+  \caption{Simple program that uses a calling convention not relyant
+    on the stack}
+  \label{fig:prog-f1}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{lstlisting}
+f2: push 1
+    fetch $r_1$ $r_\adv$
+    scall $r_1$($[]$,$[]$)
+    pop $r_1$
+    assert $r_1$ 1
+    halt
+\end{lstlisting}
+  \caption{Simple program that uses a stack-based calling convention.}
+  \label{fig:prog-f2}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{lstlisting}
+f3: push 1
+    fetch $r_1$ $\adv$
+    scall $r_1$($[]$,$[r_1]$)
+    pop $r_2$
+    assert $r_2$ 1
+    push 2
+    scall $r_1$($[]$,$[]$)
+    halt
+\end{lstlisting}
+  \caption{Simple program that relies on control-flow integrity.}
+  \label{fig:prog-f3}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \begin{lstlisting}
+g1:  malloc $r_2$ 1
+     store $r_2$ 0
+     move $\pcreg$ $r_3$
+     lea $r_3$ $\var{offset}$
+     crtcls $[(x, r_2)]$ $r_3$
+     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
+     jmp $r_0$
+f4:  reqglob $r_1$
+     prepstack $r_\stk$
+     store $x$ 0
+     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
+     store $x$ 1
+     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
+     load $r_1$ x
+     assert $r_1$ 1
+     mclear $r_\stk$
+     rclear $\RegName \setminus \{r_0,\pcreg\}$
+     jmp $r_0$
+\end{lstlisting}
+  \caption{A program that relies on well-bracketedness of
+    \texttt{scall}s to function correctly. $\var{offset}$ is the
+    offset to $f4$.}
+  \label{fig:prog-g1}
+\end{figure}
+
 \begin{itemize}

### 2017-02-09 local-capability-machine.tex
Some work on the capability machine section.
typo in awkward example

@@
 1g:  jmp $r_0$
-f4:  reqglob $r1$
+f4:  reqglob $r_1$
      prepstack $r_\stk$

### 2017-02-09 csf-2017/paper.tex
Some more discussion of the capability machine.

--- 2017-02-09 csf-2017/paper.tex
@@
 
+We define the syntax of our capability machine in \figurename~\ref{fig:syntax}. We
+assume an infinite set of addresses $\Addrs$ and define machine words as either
+integers or capabilities of the form $((\perm,\gl),\start,\addrend,\addr)$. Such
+a capability represents the authority to execute permissions $\perm$ on the
+memory array $[\start,\addrend]$, together with a current address $\addr$ and a
+tag $\gl$ which indicates whether the capability is global or local. The
+available permissions and their ordering are depicted in
+\figurename~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
+readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
+read/write/execute ($\rwx$) permissions. Additionally, there are three special
+read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter
+($\entry$) permissions which we will explain below.
+
+We assume a set of register names $\RegName$. We define register files $\reg$ as
+functions mapping register names to words and memories as functions that map
+addresses to words. The state of the entire machine is represented as a
+configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
+and a register file, or a failed or halted state, where the latter keeps hold
+of the final state of memory. We also define memory segments $\ms$:
+parts of memory represented as partial functions mapping addresses to words.
+
+The instruction set of our machine is rather basic. Instructions $i$ include
+relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
+($\move{}{}$, copies words between registers) instructions. Also familiar are
+load and store instructions for reading from and writing to memory ($\load{}{}$
+and $\store{}{}$) and an arithmetic addition operator ($\plus{}{}{}$). There are
+three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
+address), $\restrict{}{}$ (modifies the permission and local/global flag) and
+$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
+instructions take care that the resulting capability always carries less
+authority than the original (e.g. restrict will only weaken a permission).
+Finally, there is an instruction $\isptr{}{}$ for testing whether a word is a
+capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$,
+$\gete{}{}$ and $\geta{}{}$ for obtaining respectively the permissions,
+local/global flag, base, end and current address of a capability.
+
 \begin{figure*}
@@
     \begin{cases}
-      \sem{\decode(\memheap(\addr))}(\Phi) & \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ } \\
+      \sem{\decode(n)}(\Phi) & \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }\\
       \failed                                 & \text{otherwise}
     \end{cases}
-  \end{equation*}%
+  \end{equation*}\vspace{-.3cm}%
   \begin{align*}
@@
 
-  \begin{tabular}{|c|p{3.7cm}|p{10.5cm}|}
+  \begin{tabular}{|c|p{3.8cm}|p{10.4cm}|}
     \hline
@@
     \hline
-    $\jmp{\lv}$&$\begin{multlined}
-                   \updateReg{\pcreg}{\\\updatePcPerm{\memreg(\lv)}}
-                 \end{multlined}$&\\
-    \hline
     $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$  and  $\start \leq \addr \leq \addrend$\\
-    \hline
-    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and  $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
-                                                                                       and if  $\var{w} = ((\_,\local),\_,\_,\_)$, then  $\perm \in \{\rwlx,\readwritel \}$\\
     \hline
@@
     \hline
+    $\jmp{\lv}$&$\begin{multlined}
+                   \updateReg{\pcreg}{\\\updatePcPerm{\memreg(\lv)}}
+                 \end{multlined}$&\\
+    \hline
+    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and  $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
+                                                                                       and if  $\var{w} = ((\_,\local),\_,\_,\_)$, then  $\perm \in \{\rwlx,\readwritel \}$\\
+    \hline
     \multicolumn{3}{|c|}{$\cdots$}\\
@@
 
-We define the syntax of our capability machine in \figurename~\ref{fig:syntax}. We
-assume an infinite set of addresses $\Addrs$ and define machine words as either
-integers or capabilities of the form $((\perm,\gl),\start,\addrend,\addr)$. Such
-a capability represents the authority to execute permissions $\perm$ on the
-memory array $[\start,\addrend]$, together with a current address $\addr$ and a
-tag $\gl$ which indicates whether the capability is global or local. The
-available permissions and their ordering are depicted in
-\figurename~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
-readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
-read/write/execute ($\rwx$) permissions. Additionally, there are three special
-read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter
-($\entry$) permissions which we will explain below.
-
-We assume a set of register names $\RegName$. We define register files $\reg$ as
-functions mapping register names to words and memories as functions that map
-addresses to words. The state of the entire machine is represented as a
-configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
-and a register file, or a failed or halted state, where the latter keeps hold
-of the final state of memory. We also define memory segments $\ms$:
-parts of memory represented as partial functions mapping addresses to words.
-
-The instruction set of our machine is rather basic.  Instructions $i$ include
-jumps ($\jmp{}$), conditional jumps ($\jnz{}{}$), a move instruction that copies
-a word from one register to another ($\move{}{}$), load and store instructions
-for reading from and writing to memory ($\load{}{}$ and $\store{}{}$)
+Figure~\ref{fig:op-sem} shows the operational semantics of our capability
+machine, or at least an excerpt containing a number of representative
+instructions. Essentially, a configuration $\Phi$ will either decode and execute
+the instruction at $\memreg{\pcreg}$ if it is executable and the current address
+is in the valid range or fail otherwise. The table in the figure shows for a
+given instruction $i$ the result of executing it in configuration $\Phi$. The
+$\fail$ and $\halt$ instructions simply fail and halt. The $\move{}{}$
+instruction simply modifies the register file as requested and updates the
+$\pcreg$ to the next instruction using the meta-function $\stdUpdatePc{}$.
+
+The $\load{}{}$ instruction loads the contents of the requested memory location
+into the appropriate register, but only if the capability has appropriate
+authority (i.e. read permission and an appropriate range). The $\restrict{}{}$
+instruction will update a capability's permissions and global/local status in
+the register file, but requires that the new permissions are weaker than the
+original and does not turn local capabilities into global ones. The $\geta{}{}$
+instruction queries the current address of a capability and stores it in another
+register.
+
+The $\jmp{}$ instruction updates the program counter to a requested location,
+but it is complicated by the presence of \emph{enter capabilities}, modeled
+after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
+cannot be used to read, write or execute and their current address and range
+cannot be modified. They can only be used to jump to, but when that happens,
+their permission is updated to $\exec$. This special behavior makes them
+suitable for representing a kind of closures: an opaque package containing a
+piece of code together with local encapsulated state. The idea is that one can
+build an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where the
+range $[\start,\addr-1]$ is used for storing the closure's local state (data or
+capabilities) and the range $[\addr,\addrend]$ can be used for storing
+instructions. The instructions and data are opaque to an adversary holding $c$
+but when $c$ jumped to, the data becomes available to the closure's code through
+the updated version of $c$ that is now in the $\pcreg$\footnote{Note that
+  $\pcreg$ can be used to $\load{}{}$ from, like any other register.}.
+
+The final semantics we discuss is that of $\store{}{}$. As expected, it updates
+the memory to the requested value if the capability has write authority for the
+requested location. However, the instruction is complicated by the presence of
+\emph{local capabilities}, modeled after the ones in
+CHERI~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
+they can only be kept in registers, i.e. they cannot be stored to memory. This
+means that local capabilities can be \emph{temporarily} given to an adversary,
+for the duration of an invocation: if we take care to erase the capability from
+the register file after control is passed back to us, we can be sure that he/she
+will not have been able to store the capability. However, there is one exception
+to the rule above: local capabilities can be stored to a part of memory for
+which we have a capability with write-local authority (i.e. permission $\rwl$ or
+$\rwlx$). This exception is intended to accomodate regions of memory that behave
+like the C stack: stashing away the contents of registers on the stack can be
+allowed, even if this includes local capabilities. As long as all capabilities
+with write-local authority are themselves local and the stack is cleared after
+control is passed back by the adversary, we will see that this does not break
+the intended behavior of local capabilities.
 

### 2017-02-09 csf-2017/paper.tex
Minor changes

--- 2017-02-09 csf-2017/paper.tex
@@
 
+\subsection{Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
 \todo[inline]{Add discussion/comparison to other LRs}
@@
 
-\subsection{Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
-
 \section{Examples}
+
 \begin{figure}[htbp]
@@
   \begin{lstlisting}
-f1: malloc $r_l$ 1
+f1: push 1
+    fetch $r_1$ $r_\adv$
+    scall $r_1$($[]$,$[]$)
+    pop $r_1$
+    assert $r_1$ 1
+    halt
+\end{lstlisting}
+  \caption{Simple program that uses a stack-based calling convention.}
+  \label{fig:prog-f1}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+\begin{lstlisting}
+f2: malloc $r_l$ 1
     store $r_l$ 1
@@
     on the stack}
-  \label{fig:prog-f1}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{lstlisting}
-f2: push 1
-    fetch $r_1$ $r_\adv$
-    scall $r_1$($[]$,$[]$)
-    pop $r_1$
-    assert $r_1$ 1
-    halt
-\end{lstlisting}
-  \caption{Simple program that uses a stack-based calling convention.}
   \label{fig:prog-f2}

### 2017-02-09 csf-2017/paper.tex
Examples, local state encapsulatoin

@@
 
-\todo[inline]{Put paragraph about memory satisfaction here}
 \begin{align*}
@@
 \section{Examples}
-
+In this section, we display how capability safety allows us to prove
+local state encapsulation and control-flow integrity properties. For
+presentation purposes, we do not formally state the lemmas we
+prove. The intere
+sted reader can find all the technical details in the
+techincal report.
+
+\subsection{Encapsulation of local state} 
+The program \texttt{f1} in \figurename~\ref{fig:prog-f1} simply pushes 1 to the
+stack, jumps to an adversary using \texttt{scall} and finally asserts
+the top value on the stack is 1. If this program is linked up
+with an arbitrary adversary that has access to malloc but otherwise only
+consists of instructions, then we can show that the assertion never
+fails. This shows that the local state of $\texttt{f1}$ is
+encapsulated.
 \begin{figure}[htbp]
@@
 
+\todo[inline]{give a taste of a proof by giving high-level proof here?
+Maybe later for one of the other examples}
+
+In \figurename~\ref{fig:prog-2}, we have \texttt{f2}. \texttt{f2} is
+very similar to \texttt{f1}, but instead of using the stack for local
+state and the stack-based calling convention, it allocates new memory
+for the local state and use a calling convention based on activation
+record. Despite the different approach to handle local state and
+calls, we are still able to show a similar property to that of
+\texttt{f1}. This illustrates that the logical relation is not
+specific to the stack-based calling convention as it can be used to
+reason about programs that follow a different calling convention.
 \begin{figure}[htbp]
@@
 \end{figure}
+
+\subsection{Well-bracketed control-flow}
 

### 2017-02-09 csf-2017/paper.tex
changed f1 and f2 text in example section

--- 2017-02-09 csf-2017/paper.tex
@@
 \section{Examples}
-
+In this section, we display how capability safety allows us to prove
+local state encapsulation and control-flow integrity properties. For
+presentation purposes, we do not formally state the lemmas we
+prove. The interested reader can find all the technical details in the
+techincal report.
+
+\subsection{Encapsulation of local state}
+The capability machine provides encapsulation of local state. The program \texttt{f1} in \figurename~\ref{fig:prog-f1} and the program \texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.
+
+The programs \texttt{f1} and \texttt{f2} are very similar. They both
+store some local data, call an untrusted piece of code, and finally
+make sure that their local state is unchanged. They differ in the way
+they do this. \texttt{f1} assumes it has access to a stack capability
+and uses the stack to store its local state. It also uses the
+stack-based calling convention to call the untrusted code. On the
+other hand, \texttt{f2} allocates a new piece of memory for its local
+state and uses an activation-record based calling convention
+(described in the technical report) to run the adveserial code.
+
+For both programs, we can prove that if they are linked up with an
+adversary that is allowed to allocate new memory but otherwise has no
+capabilities, then executing the program, the assertion will never
+fail.
+We can only prove this, because the capability machine encapsulates
+the local state of the two programs. The two examples also illustrate
+the versatility of the logical relation. The two examples use
+different calling convention, but the logical relation is not specific
+to any calling convention, so we can use it to reason about both
+programs. In order to reason about the unknown untrusted part, we use
+the Fundamental Theorem to argue that it must be capability safe.
 \begin{figure}[htbp]
@@
 f1: push 1
-    fetch $r_1$ $r_\adv$
+    fetch $r_1$ $\adv$
     scall $r_1$($[]$,$[]$)
@@
 \end{figure}
+
+\todo[inline]{give a taste of a proof by giving high-level proof here?
+Maybe later for one of the other examples}
 
@@
 
+\subsection{Well-bracketed control-flow}
+
+The program $f3$ in \figurename~\ref{fig:prog-f3}
 \begin{figure}[htbp]

### 2017-02-09 csf-2017/paper.tex
f3 text

@@
 the Fundamental Theorem to argue that it must be capability safe.
+
+\todo[inline]{give a more detailed high-level taste of a proof here? Maybe later for one of the other examples?}
+
 \begin{figure}[htbp]
@@
 \end{figure}
-
-\todo[inline]{give a taste of a proof by giving high-level proof here?
-Maybe later for one of the other examples}
 
@@
 \subsection{Well-bracketed control-flow} 
-
-The program $f3$ in \figurename~\ref{fig:prog-f3}
+Using the stack-based calling convention of \texttt{scall}, we get
+well-bracketed control-flow. To illustrate this, we look at
+two example programs \texttt{f3} in \figurename~\ref{fig:prog-f3} and
+\texttt{g1} in \figurename~\ref{fig:prog-g1}.
+
+In \texttt{f3} there are two calls to an adversary and in order for
+the assertion in the middle to succeed, they need to be
+well-bracketed. If the adversary is able to store the protected return
+pointer, so it is available for the second call, then it can be used
+in a context where \texttt{f3} has 2 on top of its stack which will cause the
+assertion to fail. The protected return pointer is passed as a local
+capability, so the only means the adversary has to store it is the
+stack capability. The stack capability itself is a local capability,
+so the adversary can only store it on the stack. This means that after
+the adversary jumps to the protected return pointer, \texttt{f3} can
+revoke it (by overwritting it which happens as part of the
+\texttt{scall}). In other words, the adversary cannot store the
+protected return pointer for the second call, and it is possible to
+prove that the assertion never fails.
 \begin{figure}[htbp]
@@
 \end{figure}
+
+Our final example is \texttt{g1} which is a faithful translation of
+what is known as the ``awkward'' example.
+\todo[inline]{insert references}
 

### 2017-02-09 csf-2017/paper.tex
lr pass

@@
 
+\subsubsection{World satisfaction}
+A memory satisfies a world, written $\memSat{\ms}{W}$, 
+if it can be partitioned into disjoint parts such that each part is
+in the interpretation of an active (permanent or temporary) region. 
 \begin{align*}
@@
 \end{align*}
-
-A memory satisfies a world, written $\memSat{\ms}{W}$, 
-if it can be partitioned into disjoint parts such that each part is
-in the interpretation of an active (permanent or temporary) region.
 
@@
 
-The \emph{expression relation}, $\stder$, defines what words are
-capability safe when used for the program counter. A word is
-capability safe as a program counter when it can be plugged into a
+The \emph{expression relation}, $\stder$, defines when a word is
+capability safe as a program counter. That is the case when the word can be plugged into a
 capability safe register file (i.e., a register file in $\stdrr$) and
@@
 anywhere in memory, so they must be capability safe as a program
-counter no matter where they point in memory. In contrast, enter
-\lars{``capability safe as a program counter'' is a bit hard to
-  understand - why not say that this is a capability one can jump to ?}
-\lau{ When jumping to a capability, it becomes the program counter.
-A program counter is capability safe when it is the $\stder$ relation.
-When I say ``capability safe as a program counter'' I take it to mean being in the $\stder$-relation.
-}
+counter (in the $\stder$-relation) no matter where they point in memory. In contrast, enter
 capabilities are opaque and must be used from the address they point
@@
 read/execute.
-In both conditions, it must be safe to jump to the capability. This
+In both cases, it must be safe to jump to the capability. This
 means that the capability must be capability safe as a
-program counter in certain future worlds. If the capability is global,
-then it should be capability safe as a program counter in any private
-future world. This means that it can only depend on permanent regions
-because temporary regions may be revoked in a private future
-world. On the other hand, local capabilities must only be capability
-safe in public future worlds because they may depend on temporary
+program counter in a future world. If the capability is global,
+then it should only depend on permanent regions and therefore
+we require that it should be capability safe as a program counter in any \emph{private}
+future world (recall that temporary regions may be revoked in a private future
+world). On the other hand, a local capability should be capability
+safe in any \emph{public} future world because it may depend on temporary
 regions, so we cannot expect them to function correctly as program
@@
 With the logical relation defined, we can now state a strong theorem
-about the safety properties of the capability machine. Namely that any
-capability that does not grant unreasonable authority is capability
+about the safety properties of the capability machine:
+any capability that only grants reasonable authority is capability
 safe as a program counter.
@@
 The Fundamental Theorem allows us to prove capability safety for
-untrusted programs. If we can show that the authority the capability
-for the program grants is reasonable, then it is safe to
-execute. Intuitively this is because the logical relation captures the
-safe behaviour of the capability machine that do not allow you to
-produce something capability \emph{unsafe} from something capability
-safe. The Fundamental Theorem also serves as a sanity check of the
-logical relation as it makes sure that a large class of capabilities
+untrusted programs: if the authority granted by
+the capabilty for the untrustred program is reasonable, then it is safe to
+execute. Intuitively, this is because our logical relation captures the
+safe behaviours of the capability machine that do not allow you to
+produce something \emph{unsafe} from something safe. 
+The Fundamental Theorem also serves as a sanity check of the
+logical relation as it guarantees that a large class of capabilities
 actually inhabits it.
@@
 
-\bibliographystyle{IEEEtran}
+%\bibliographystyle{IEEEtran}
+\bibliographystyle{plain}
 \bibliography{IEEEabrv,references}

### 2017-02-09 csf-2017/paper.tex
lr

@@
 
-\subsection{Discussion/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
-\todo[inline]{Add discussion/comparison to other LRs}
+\subsection{Discussion}
+%/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
+%\todo[inline]{Add discussion/comparison to other LRs}
 % Comparisson of LR to other work:
@@
 %   capabilities can handle.
+
+\todo[inline]{Maybe move this subsection to related work}
+
+As already mentioned, our the state transition systems used in our
+logical relation are inspired by earlier work on logical relations for
+high-level
+languages~\cite{Ahmed:popl09, Dreyer:jfp12,Devriese:2016ObjCap}.
+The idea of public and private transitions comes from 
+Dreyer et. al.~\cite{Dreyer:jfp12}, who gave a logical relations model
+of a high-level lambda calculus. In \emph{loc. cit.}
+public transitions were used to 
+capture the \emph{extensional} behavior of a program expression
+(and a context), whereas private transitions were used to capture
+the internal behaviour of a computation using local state.
+Technically, the logical relation of Dreyer et. al.\ was
+formulated using biorthogonal closure, and the future world
+relation for continuations was required to be a public future world
+relation. In other words, the choice between private or public future
+rested on the ``structure'' of the programming language. 
+In contrast, 
+for our low-level, unstructured, assembly language, we
+we instead decide on whether to use a private or public future 
+world depending on whether the capability is global or local. 
+Another interesting technical difference is that in our low-level 
+language we do not have a fixed calling convention and any register
+can serve as a continuation, and therefore we do not use biorthogonal
+closure but just model return pointers as continuations using a single
+orthogonal closure.
+
+\lars{don't know if we need the following paragraph:}
+(Even though 
+Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also
+considered a logical
+relation involving an assembly language, they still used
+biorthogonal closure but that was because their logical relation
+was a relation between a high-level language and an assembly
+language,  and thus the assembly language in the relation essentially
+inherited structure from the high-level language.)
 

### 2017-02-09 csf-2017/paper.tex
some text for g1

--- 2017-02-09 csf-2017/paper.tex
@@
 
-Our final example is \texttt{g1} which is a faithful translation of
-what is known as the ``awkward'' example.
-\todo[inline]{insert references}
+The final example in \figurename~\ref{fig:prog-g1} consists of two
+part: \texttt{g1} creates a closure with one variable $x$ in its
+environment and \texttt{f4} as the code. When the closure is called,
+it expects to be called with a capability for a callback and a stack
+capability.
+
+For the property we show for \texttt{g1}, the control is inverted
+compared to the other examples.
 

### 2017-02-09 csf-2017/paper.tex
Start on section about stack and return pointer management discipline.

--- 2017-02-09 csf-2017/paper.tex
@@
 
+% Contrary to more traditional assembly languages, capability machines do not
+% allow literal addresses to appear as part of instructions. This means that we
+
 Explain our model of linking
+\dominique{not sure what we want to say about it here...}
 
 \section{Stack and return pointer management using local capabilities}
+
+One of the contributions in this paper is a demonstration that local
+capabilities on a capability machine can be used in a stack and return pointer
+management scheme that enforces control flow integrity in a way that is provably
+watertight, potentially efficient, does not rely on a trusted central stack
+manager, and with support for exposing higher-order interfaces to an adversary.
+In this section, we explain this discipline. We start by explaining the
+high-level approach, the different issues that need to be taken into account,
+starting with first-order interfaces and moving onto a higher-order setting
+after that. After that, we define a number of reusable macro-instructions that
+can be used to conveniently apply the proposed discipline in subsequent
+examples.
+
+The basic idea of our approach is actually rather simple: we stick to a single,
+rather standard C stack and register-passed return pointers in much the same way
+as a standard C ABI. However, to prevent various ways of misusing this basic
+scheme, we put local capabilities to work and need to perform a number of
+not-always-obvious checks and safety measures.
+
+\paragraph{Invoking the adversary}
+* stack and return pointer as local caps
+* clearing registers
+* clearing the stack
+
+\paragraph{Returning from the adversary}
+* return pointer as enter pointer to activation record on the stack.
+* activation record contains previous pc and stack pointer
+
+\paragraph{Invoking an adversary callback}
+* must be global
+
+\paragraph{Having a callback invoked by the adversary}
+* checking the received stack pointer
+
 \begin{itemize}

### 2017-02-09 csf-2017/paper.tex
more g1

--- 2017-02-09 csf-2017/paper.tex
@@
 The final example in \figurename~\ref{fig:prog-g1} consists of two
-part: \texttt{g1} creates a closure with one variable $x$ in its
-environment and \texttt{f4} as the code. When the closure is called,
-it expects to be called with a capability for a callback and a stack
-capability.
+part: \texttt{g1} is a closure generation program with one variable $x$ in its
+environment and \texttt{f4} as the code.
+Like in the previous example, the assertion can fail if the adversary can capture the protected return pointer of the second call, invoke the closure anew and use the captured protected return pointer in the first call when the variable $x$ has the value 0.
+
+In the other examples, the known program started the execution and called an unknown piece of code.
+In this example, the control is inverted, so an unknown piece of code gets an enter capability for the closure creator.
 
@@
      scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
-     load $r_1$ x
+     load $r_1$ $x$
      assert $r_1$ 1

### 2017-02-09 csf-2017/paper.tex

--- 2017-02-09 csf-2017/paper.tex
@@
 
-Our final example is \texttt{g1} which is a faithful translation of
-what is known as the ``awkward'' example.
-\todo[inline]{insert references}
+The final example in \figurename~\ref{fig:prog-g1} consists of two
+part: \texttt{g1} is a closure generation program with one variable $x$ in its
+environment and \texttt{f4} as the code.
+Like in the previous example, the assertion can fail if the adversary can capture the protected return pointer of the second call, invoke the closure anew and use the captured protected return pointer in the first call when the variable $x$ has the value 0.
+
+In the other examples, the known program started the execution and called an unknown piece of code.
+In this example, the control is inverted, so an unknown piece of code gets an enter capability for the closure creator.
+
+For the property we show for \texttt{g1}, the control is inverted
+compared to the other examples.
 

### 2017-02-10 csf-2017/paper.tex
minor

@@
 For reasons of space, some details and all proofs have been omitted;
-please refer to the technical appendix~\cite{} for those.
+please refer to the technical appendix~\cite{XXX} for those.
 
@@
 explained when needed. The precise definitions can be found in the
-technical report.\todo[inline]{add reference to technical report}
+technical report~\cite{XXX}.
 
@@
 
-\lars{don't know if we need the following paragraph:}
-(Even though 
+\lars{don't know yet whether we need something like the following paragraph:}
+Even though 
 Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also
@@
 language,  and thus the assembly language in the relation essentially
-inherited structure from the high-level language.)
+inherited structure from the high-level language.
 
@@
 presentation purposes, we do not formally state the lemmas we
-prove. The interested reader can find all the technical details in the
+prove. 
+\lars{I think it would be good if we can state one of them,
+  possibly in shortened form, just to make it clear that something is
+  actually stated and proved formally, and make a link to the LR}
+The interested reader can find all the technical details in the
 techincal report.

### 2017-02-10 csf-2017/paper.tex
added copy of f1 lemma

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 
+\begin{lemma}[Correctness lemma for \texttt{f1}]
+  \label{lem:correctness-f1}
+  Let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
+    \reg & \in \Regs \\
+    m & \defeq \hs_{f1} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus 
+        \ms_{\malloc} \uplus 
+        \ms_{\var{stk}} \uplus
+        \ms_{\var{frame}} 
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
+  \end{itemize}
+  where 
+  \begin{align*}
+    &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link+1]\\
+    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
+  \item $\ms_\flag = [\flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+  \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
 \begin{figure}[htbp]

### 2017-02-10 csf-2017/paper.tex
suggestion for paper edition of f1 lemma
short text about f1 lemma proof

@@
 \subsection{Encapsulation of local state}
-The capability machine provides encapsulation of local state. The program \texttt{f1} in \figurename~\ref{fig:prog-f1} and the program \texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.
+The capability machine provides encapsulation of local state. The
+program \texttt{f1} in \figurename~\ref{fig:prog-f1} and the program
+\texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.
 
@@
 For both programs, we can prove that if they are linked up with an
-adversary that is allowed to allocate new memory but otherwise has no
+adversary, $\adv$, that is allowed to allocate new memory but otherwise has no
 capabilities, then executing the program, the assertion will never
@@
 to any calling convention, so we can use it to reason about both
-programs. In order to reason about the unknown untrusted part, we use
-the Fundamental Theorem to argue that it must be capability safe.
+programs.
 
 \todo[inline]{give a more detailed high-level taste of a proof here? Maybe later for one of the other examples?}
-
 \begin{lemma}[Correctness lemma for \texttt{f1}]
+\todo[inline]{The lemma copied directly from the TR. To be deleted later!}
   \label{lem:correctness-f1}
@@
 
+In order to formulate a lemmas about \texttt{f1} and \texttt{f2}, we need a way to observe whether the assertion fails.
+To this end, we assum access to a flag which will be set by the assertion if it fails.
+A slightly simplified version of the correctness lemma for \texttt{f1} is:
+\begin{lemma}
+  \label{lem:correctness-f1}
+  Let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\dots) \\
+    c_{f1} & \defeq ((\rwx,\glob),\dots) \\
+    c_\malloc & \defeq ((\entry,\glob),\dots) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\dots) \\
+    c_\link & \defeq ((\readonly,\glob),\dots)\\
+    \reg & \in \Regs \\
+    m & \defeq
+        \begin{aligned}[t]
+          &\hs_{f1} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus \\
+        &\ms_{\malloc} \uplus 
+        \ms_{\var{stk}} \uplus
+        \ms_{\var{frame}}
+      \end{aligned}
+  \end{align*}
+  where each of the capability has an appropriate range of authority and pointer and
+  \begin{itemize}
+  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
+  \item $\ms_\flag(\flag) = 0$
+  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
+  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[*] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
+In order for prove Lemma~\ref{lem:correctness-f1}, it suffices to show
+that the start configuration is capability safe (in the $\observations$ relation) for a
+world with a permanent region that requires the assertion flag to be
+0. This is done using an anti-reduction lemma which states that for a
+configuration to be capability safe, it is sufficient to
+show that it is in it at a later point of the execution. To reason
+about \texttt{scall}, it suffices to show that the configuration just
+after we jumped is capability safe and that the configuration just
+after \texttt{scall} is done cleaning up is capability safe.
+We use the Fundamental Theorem to reason about the unknown code, but
+notice that the Fundamental Theorem says nothing about enter
+capabilities and that the adversary capability is an enter
+capability. Luckily the enter capability becomes a read/execute
+capability when we jump and the Fundamental Theorem can be used to
+show that this capability is capability safe as a program counter.
+
 \begin{figure}[htbp]

### 2017-02-10 csf-2017/paper.tex
added lemma for f2
Some work on explaining the stack and return pointer management discipline.

--- 2017-02-10 csf-2017/paper.tex
@@
 % One argument
-\newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
+\newcommand{\oneinstr}[2]{
+  \ifthenelse{\equal{#2}{}}
+  {\zinstr{#1}}
+  {\zinstr{#1} \; #2}
+}
 \newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
 % Two arguments
-\newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
+\newcommand{\twoinstr}[3]{
+  \ifthenelse{\equal{#2#3}{}}
+  {\zinstr{#1}}
+  {\zinstr{#1} \; #2 \; #3}
+}
 \newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
@@
 % Three arguments
-\newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
+\newcommand{\threeinstr}[4]{
+  \ifthenelse{\equal{#2#3#4}{}}
+  {\zinstr{#1}}
+  {\zinstr{#1} \; #2 \; #3 \; #4}
+}
 \newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
@@
 three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
-address), $\restrict{}{}$ (modifies the permission and local/global flag) and
+address), $\restricttwo{}{}$ (modifies the permission and local/global flag) and
 $\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
@@
 into the appropriate register, but only if the capability has appropriate
-authority (i.e. read permission and an appropriate range). The $\restrict{}{}$
+authority (i.e. read permission and an appropriate range). The $\restricttwo{}{}$
 instruction will update a capability's permissions and global/local status in
@@
 
+Explain malloc
+\dominique{not sure what we want to say about it here either...}
+
 \section{Stack and return pointer management using local capabilities}
@@
 
-The basic idea of our approach is actually rather simple: we stick to a single,
-rather standard C stack and register-passed return pointers in much the same way
-as a standard C ABI. However, to prevent various ways of misusing this basic
-scheme, we put local capabilities to work and need to perform a number of
-not-always-obvious checks and safety measures.
+The basic idea of our approach is quite simple: we stick to a single, rather
+standard C stack and register-passed return pointers in much the same way as a
+standard C ABI. However, to prevent various ways of misusing this basic scheme,
+we put local capabilities to work and take a number of not-always-obvious safety
+measures. In the next paragraphs, we will explain the issues to be considered in
+all the relevant situations: when (1) starting our program, (2) invoking the
+adversary, (3) returning from the adversary, (4) invoking an adversary callback
+and (5) having a callback invoked by the adversary.
+
+\dominique{Add a summary table somewhere?}
+
+\paragraph{Program startup}
+When our code starts executing, we assume that the language runtime initializes
+the memory as follows: a contiguous array of memory is reserved for the stack,
+for which we receive a stack pointer in a register that we reserve for this
+purpose \dominique{give this register a special name like sp?}. This stack
+pointer must be local and it must carry $\rwlx$ permission. It is crucial for
+our approach that the stack is the only part of memory for which there will ever
+be capabilities with $\rwlx$ or $\rwl$ capabilities. We expect the language
+runtime to enforce this by not handing out such capabilities for other parts of
+memory in any situation (i.e. not through malloc, not during loading or linking
+etc.). Additionally, in our examples, we typically also assume some static parts
+of memory where instructions are stored or static data. Another part of memory
+(called the heap) is initially governed by malloc and at program startup, no
+other code has capabilities for this memory. Malloc will hand out capabilities
+for allocated regions as requested, but it will never hand out capabilities with
+write-local capability (i.e. no $\rwlx$ or $\rwl$ permissions). For simplicity,
+we assume that memory allocated through malloc cannot be freed.
 
 \paragraph{Invoking the adversary}
-* stack and return pointer as local caps
-* clearing registers
-* clearing the stack
+During execution, we use the stack as normal: we may push values onto it and pop
+them back off (e.g. for register spilling etc.). When we invoke the adversary,
+we need to take great care to only provide him with the authority we want him to
+have. Specifically, we clear all the registers, except those that contain
+arguments the adversary is intended to have access to. We leave a copy of the
+stack pointer in the appropriate register, but only after we have used the
+$\subseg{}{}{}$ to shrink its authority to just the part of the stack that we
+are not using. In one of the registers, we will also provide a return pointer.
+We will go into constructing the return pointer further below.
+
+It is already important to note that the return pointer
+will be provided to the adversary as local capabilities. If the adversary were
+able to store away the return pointer in a global data structure, then he would
+be able to jump to it in circumstances where this should not be possible: he
+may, for example, store the return pointer, legally jump to it a first time,
+wait to be invoked again and then jump to the old return pointer a second time,
+rather than the new return pointer received for the new invocation. Similarly,
+he may store the return pointer, invoke a function in the trusted code, wait to
+be invoked again and then jump to the stored return pointer rather than the new
+return pointer for the second invocation.
+
+Because the return pointer is local, the adversary is only able to store it
+through write-local capabilities, which (because of our assumptions above)
+means: on the stack. However, this makes it crucial that the stack pointer is
+also local: if the adversary were able to store the stack pointer in a global
+data structure, he/she can store return pointers on the stack, and play similar
+mischief as mentioned above. Storing away stack pointers for use during future
+invocations would also be dangerous for other reasons: during such a future
+invocation, a larger part of the stack may be in use by the trusted code, which
+the adversary may still have access to through the old stack pointer.
+
+A final measure we need to take is that before invocation of the adversary, we
+need to clear (i.e. overwrite with zeros) the part of the stack that we are not
+using.  Otherwise, that part of the stack may still contain data and
+capabilities from previously executed code, including local capabilities since
+they are allowed on the stack.  The adversary would be able to take advantage of
+this by using it to store either stack pointers or return pointers from previous
+invocations and play the same tricks as above.
 
 \paragraph{Returning from the adversary}
-* return pointer as enter pointer to activation record on the stack.
-* activation record contains previous pc and stack pointer
+So return pointers must be passed as local capabilities. But what should their
+permissions be, what memory should they point to and what should that memory
+(let's call it the activation record) contain? Let us answer the last question
+first by considering what should happen when the adversary jumps to a return
+pointer and what capabilities we need for achieving that. Essentially, the
+program counter should be restored to the code after the call into the
+adversary, so the activation record should store this program counter.
+Additionally, the stack pointer should also be restored to its original value.
+Since the adversary has a more restricted authority over the stack than the code
+making the call, we cannot hope to reconstruct the return pointer from the stack
+pointer owned by the adversary. Instead, the old stack pointer should also be
+stored as part of the activation record.
+
+\begin{figure}
+  \begin{tikzpicture}
+    \draw[fill=gray!20,draw=none] (0,4) rectangle (4,3.5);
+    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (4,1);
+    \draw (0,0.5) -- (0,4);
+    \draw (4,0.5) -- (4,4);
+    \draw (0,3.5) rectangle (4,3) node[pos=.5] {\footnotesize previous program counter};
+    \draw (0,3) rectangle (4,2.5) node[pos=.5] {\footnotesize previous stack pointer};
+    \draw (0,2.5) rectangle (4,1) node[pos=.5] {\footnotesize restore instructions};
+    \draw (-2,2.25) node (rp) {return pointer};
+    \draw[->] (rp.east) -- (-.2,2.25);
+  \end{tikzpicture}
+
+  \caption{Structure of an activation record}
+  \label{fig:activ-rec-struct}
+\end{figure}
+
+It's important that neither of these capabilities are accessible to the
+adversary. In other words, we need to provide the adversary a capability that he
+can jump to but not read from, i.e. an enter capability. To make this work, we
+construct the activation record as depicted in Figure~\ref{fig:activ-rec-struct}.
+
+* return pointer as enter pointer to activation
+record on the stack. * activation record contains previous pc and stack pointer
 

### 2017-02-10 csf-2017/paper.tex

@@
 \end{figure}
-\clearpage
 }
@@
 
+In order to formulate a lemmas about \texttt{f1} and \texttt{f2}, we need a way to observe whether the assertion fails.
+To this end, we assum access to a flag which will be set by the assertion if it fails.
+A slightly simplified version of the correctness lemma for \texttt{f1} is:
+\begin{lemma}
+  \label{lem:correctness-f1}
+  Let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\dots) \\
+    c_{f1} & \defeq ((\rwx,\glob),\dots) \\
+    c_\malloc & \defeq ((\entry,\glob),\dots) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\dots) \\
+    c_\link & \defeq ((\readonly,\glob),\dots)\\
+    \reg & \in \Regs \\
+    m & \defeq
+        \begin{aligned}[t]
+          &\hs_{f1} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus \\
+        &\ms_{\malloc} \uplus 
+        \ms_{\var{stk}} \uplus
+        \ms_{\var{frame}}
+      \end{aligned}
+  \end{align*}
+  where each of the capability has an appropriate range of authority and pointer and
+  \begin{itemize}
+  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
+  \item $\ms_\flag(\flag) = 0$
+  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
+  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[*] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
+In order for prove Lemma~\ref{lem:correctness-f1}, it suffices to show
+that the start configuration is capability safe (in the $\observations$ relation) for a
+world with a permanent region that requires the assertion flag to be
+0. This is done using an anti-reduction lemma which states that for a
+configuration to be capability safe, it is sufficient to
+show that it is in it at a later point of the execution. To reason
+about \texttt{scall}, it suffices to show that the configuration just
+after we jumped is capability safe and that the configuration just
+after \texttt{scall} is done cleaning up is capability safe.
+We use the Fundamental Theorem to reason about the unknown code, but
+notice that the Fundamental Theorem says nothing about enter
+capabilities and that the adversary capability is an enter
+capability. Luckily the enter capability becomes a read/execute
+capability when we jump and the Fundamental Theorem can be used to
+show that this capability is capability safe as a program counter.
+
+For \texttt{f2} we state a similar lemma:
+\begin{lemma}
+  \label{lem:correctness-f2}
+  Making similar assumptions about capabilities and linking as in
+  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
+  if $(\reg\update{\pcreg}{c_{f2}},m) \step[n] (\halted,m')$, then $m'(\flag) = 0$.
+\end{lemma}
+
 \begin{figure}[htbp]

### 2017-02-10 csf-2017/paper.tex
Shrink figure a bit.

@@
 \begin{figure}
-  \begin{tikzpicture}
-    \draw[fill=gray!20,draw=none] (0,4) rectangle (4,3.5);
-    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (4,1);
+  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
+    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
+    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
     \draw (0,0.5) -- (0,4);
-    \draw (4,0.5) -- (4,4);
-    \draw (0,3.5) rectangle (4,3) node[pos=.5] {\footnotesize previous program counter};
-    \draw (0,3) rectangle (4,2.5) node[pos=.5] {\footnotesize previous stack pointer};
-    \draw (0,2.5) rectangle (4,1) node[pos=.5] {\footnotesize restore instructions};
-    \draw (-2,2.25) node (rp) {return pointer};
+    \draw (6,0.5) -- (6,4);
+    \draw (0,3.5) rectangle (6,3) node[pos=.5] {\footnotesize previous program counter};
+    \draw (0,3) rectangle (6,2.5) node[pos=.5] {\footnotesize previous stack pointer};
+    \draw (0,2.5) rectangle (6,1) node[pos=.5] {\footnotesize restore instructions};
+    \draw (-3,2.25) node (rp) {return pointer};
     \draw[->] (rp.east) -- (-.2,2.25);

### 2017-02-10 csf-2017/paper.tex
Finish the discussion about invoking the adversary.

@@
 
-\paragraph{Program startup}
+\textbf{Program startup}
 When our code starts executing, we assume that the language runtime initializes
@@
 
-\paragraph{Invoking the adversary}
+\textbf{Invoking the adversary}
 During execution, we use the stack as normal: we may push values onto it and pop
@@
 
-\paragraph{Returning from the adversary}
+\textbf{Returning from the adversary}
 So return pointers must be passed as local capabilities. But what should their
@@
     \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
-    \draw (0,0.5) -- (0,4);
-    \draw (6,0.5) -- (6,4);
+    \draw[draw=gray!80] (0,0.5) -- (0,4);
+    \draw[draw=gray!80] (6,0.5) -- (6,4);
     \draw (0,3.5) rectangle (6,3) node[pos=.5] {\footnotesize previous program counter};
@@
     \draw (-3,2.25) node (rp) {return pointer};
-    \draw[->] (rp.east) -- (-.2,2.25);
+    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,2.25);
   \end{tikzpicture}
@@
 can jump to but not read from, i.e. an enter capability. To make this work, we
-construct the activation record as depicted in Figure~\ref{fig:activ-rec-struct}.
-
-* return pointer as enter pointer to activation
-record on the stack. * activation record contains previous pc and stack pointer
-
-\paragraph{Invoking an adversary callback}
+construct the activation record as depicted in
+Figure~\ref{fig:activ-rec-struct}. The return pointer is an $\entry$ capability
+pointing with authority over a data structure containing the previous return and
+stack pointer as well as a number of restore instructions. The current address
+points to those instructions, so that, upon invocation, they can load the old
+stack pointer and program counter into the register file through the authority
+of the pc. The return pointer is given to the adversary as an enter pointer so
+that the adversary cannot get hold of the contents of the data structure, but
+after invocation, its permission is updated to $\exec$, so that the activation
+record's contents become available to the restore instructions.
+
+The final question that remains is: where should we construct this return
+pointer? The attentive reader will see that there is only one possibility: since
+the activation record contains the old stack pointer, which is necessarily
+local, the activation record can only be constructed in a part of memory where
+we have write-local access, i.e. on the stack. Note that this means we will be
+placing and executing instructions on the stack, i.e. it will not just contain
+code pointers and data.  This means that this discipline should be combined with
+protection against stack smashing attacks (i.e. buffer overflows on the stack overwriting activation
+records' contents), but capability machines' fine-grained memory protection
+luckily is well-suited for this.
+
+\textbf{Invoking an adversary callback}
 * must be global

### 2017-02-10 csf-2017/paper.tex
Finish the discussion of the stack and return pointer management stuff.

@@
 measures. In the next paragraphs, we will explain the issues to be considered in
-all the relevant situations: when (1) starting our program, (2) invoking the
-adversary, (3) returning from the adversary, (4) invoking an adversary callback
-and (5) having a callback invoked by the adversary.
+all the relevant situations: when (1) starting our program, (2) Returning to the
+adversary, (3) invoking the
+adversary, (4) returning from the adversary, (5) invoking an adversary callback
+and (6) having a callback invoked by the adversary.
 
@@
 
-\textbf{Invoking the adversary}
-During execution, we use the stack as normal: we may push values onto it and pop
-them back off (e.g. for register spilling etc.). When we invoke the adversary,
-we need to take great care to only provide him with the authority we want him to
-have. Specifically, we clear all the registers, except those that contain
-arguments the adversary is intended to have access to. We leave a copy of the
-stack pointer in the appropriate register, but only after we have used the
-$\subseg{}{}{}$ to shrink its authority to just the part of the stack that we
-are not using. In one of the registers, we will also provide a return pointer.
-We will go into constructing the return pointer further below.
-
-It is already important to note that the return pointer
-will be provided to the adversary as local capabilities. If the adversary were
-able to store away the return pointer in a global data structure, then he would
-be able to jump to it in circumstances where this should not be possible: he
-may, for example, store the return pointer, legally jump to it a first time,
-wait to be invoked again and then jump to the old return pointer a second time,
-rather than the new return pointer received for the new invocation. Similarly,
-he may store the return pointer, invoke a function in the trusted code, wait to
-be invoked again and then jump to the stored return pointer rather than the new
-return pointer for the second invocation.
-
-Because the return pointer is local, the adversary is only able to store it
-through write-local capabilities, which (because of our assumptions above)
-means: on the stack. However, this makes it crucial that the stack pointer is
-also local: if the adversary were able to store the stack pointer in a global
-data structure, he/she can store return pointers on the stack, and play similar
-mischief as mentioned above. Storing away stack pointers for use during future
-invocations would also be dangerous for other reasons: during such a future
-invocation, a larger part of the stack may be in use by the trusted code, which
-the adversary may still have access to through the old stack pointer.
-
-A final measure we need to take is that before invocation of the adversary, we
+\textbf{Returning to the adversary} Perhaps the simplest scenario is the case of
+returning to the adversary after he/she invoked our code. In this case, we will
+have received a return pointer from him, and we just need to jump to it as
+usual. The only thing we need to take care of is to properly clear the registers
+before we do (erasing any data we do not want to pass as (part of) the return
+value) and similarly, erase any part of the stack that we have used (i.e.
+overwrite it with zeros). Note that we can use the stack as normal: we may push
+values onto it and pop them back off (e.g. for register spilling etc.).
+
+\textbf{Invoking the adversary} Slightly more complex is the case where we
+invoke the adversary. As above, we clear all the registers, except those that
+contain arguments the adversary is intended to have access to. We leave a copy
+of the stack pointer in the appropriate register, but only after we have used
+the $\subseg{}{}{}$ to shrink its authority to just the part of the stack that
+we are not using. In one of the registers, we also provide a return pointer,
+which we will say more about below.
+
+Already important is the fact that the return pointer must be provided to the
+adversary as a local capability. If not, the adversary would be able to store
+away the return pointer in a global data structure, and he would be able to jump
+to it in circumstances where this should not be possible: he may, for example,
+store the return pointer, legally jump to it a first time, wait to be invoked
+again and then jump to the old return pointer a second time, rather than the new
+return pointer received for the new invocation. Similarly, he may store the
+return pointer, invoke a function in our code, wait for us to invoke him again
+and then jump to the stored, old return pointer rather than the new one he
+received upon the second invocation.
+
+Because the return pointer is local, this is not possible: the adversary can
+only store it through write-local capabilities, which (because of our
+assumptions above) means: on the stack. However, this makes it crucial that the
+stack pointer is also local: if the adversary were able to store the stack
+pointer in a global data structure, he/she can store return pointers on the
+stack, and play similar mischief as mentioned above. Storing away stack pointers
+for use during future invocations would also be dangerous for other reasons:
+during such a future invocation, a larger part of the stack may be in use by the
+trusted code, which the adversary may still have access to through the old stack
+pointer.
+
+A final required safety measure is that before invocation of the adversary, we
 need to clear (i.e. overwrite with zeros) the part of the stack that we are not
-using.  Otherwise, that part of the stack may still contain data and
-capabilities from previously executed code, including local capabilities since
-they are allowed on the stack.  The adversary would be able to take advantage of
-this by using it to store either stack pointers or return pointers from previous
-invocations and play the same tricks as above.
-
-\textbf{Returning from the adversary}
-So return pointers must be passed as local capabilities. But what should their
-permissions be, what memory should they point to and what should that memory
-(let's call it the activation record) contain? Let us answer the last question
-first by considering what should happen when the adversary jumps to a return
-pointer and what capabilities we need for achieving that. Essentially, the
-program counter should be restored to the code after the call into the
-adversary, so the activation record should store this program counter.
-Additionally, the stack pointer should also be restored to its original value.
-Since the adversary has a more restricted authority over the stack than the code
-making the call, we cannot hope to reconstruct the return pointer from the stack
-pointer owned by the adversary. Instead, the old stack pointer should also be
-stored as part of the activation record.
+using. Otherwise, that part of the stack may still contain data and capabilities
+from previously executed code, including local capabilities since they are
+allowed on the stack. The adversary could take advantage of this by using it to
+store either stack pointers or return pointers from previous invocations and
+play the same tricks as above. Clearing the entire unused part of the stack is
+the most problematic aspect of our discipline in terms of efficiency, since it
+may consist of a large part of memory. We believe that capability machines like
+CHERI should consider to provide special support for this requirement, by
+implementing an optimized instruction for clearing (i.e. overwriting with zeros)
+a large block of memory. We have no experience in processor design, but we
+imagine this might be optimized as a vector instruction or by having the
+processor keep track of a range of memory for which loads should return zero
+irrespective of the actual contents of the memory.
+
+\textbf{Returning from the adversary} So return pointers must be passed as local
+capabilities. But what should their permissions be, what memory should they
+point to and what should that memory (the activation record) contain? Let us
+answer the last question first by considering what should happen when the
+adversary jumps to a return pointer and what capabilities we need for achieving
+that. In that case, the program counter should be restored to point after the
+instruction that jumped into the adversary, so the activation record should
+store this old program counter. Additionally, the stack pointer should also be
+restored to its original value. Since the adversary has a more restricted
+authority over the stack than the code making the call, we cannot hope to
+reconstruct the return pointer from the stack pointer owned by the adversary.
+Instead, the old stack pointer should also be stored as part of the activation
+record.
 
@@
 It's important that neither of these capabilities are accessible to the
-adversary. In other words, we need to provide the adversary a capability that he
-can jump to but not read from, i.e. an enter capability. To make this work, we
-construct the activation record as depicted in
+adversary. In other words, the return pointer provided to the adversary must be
+a capability that he/she can jump to but not read from, i.e. an enter
+capability. To make this work, we construct the activation record as depicted in
 Figure~\ref{fig:activ-rec-struct}. The return pointer is an $\entry$ capability
-pointing with authority over a data structure containing the previous return and
-stack pointer as well as a number of restore instructions. The current address
-points to those instructions, so that, upon invocation, they can load the old
-stack pointer and program counter into the register file through the authority
-of the pc. The return pointer is given to the adversary as an enter pointer so
-that the adversary cannot get hold of the contents of the data structure, but
-after invocation, its permission is updated to $\exec$, so that the activation
-record's contents become available to the restore instructions.
-
-The final question that remains is: where should we construct this return
-pointer? The attentive reader will see that there is only one possibility: since
-the activation record contains the old stack pointer, which is necessarily
+with authority over the activation record containing the previous return and
+stack pointer as well as a number of restore instructions. The return pointer's
+current address points to those instructions, so that, upon invocation, they are
+executed and can load the old stack pointer and program counter back into the
+register file. Because the return pointer is given to the adversary as an enter
+pointer, he/she cannot get hold of the contents of the data structure, but after
+invocation, its permission is updated to $\exec$, so the activation record's
+contents become available to the restore instructions.
+
+The final question that remains is: where should we store this activation
+record? The attentive reader may already see that there is only one possibility:
+since the activation record contains the old stack pointer, which is necessarily
 local, the activation record can only be constructed in a part of memory where
@@
 placing and executing instructions on the stack, i.e. it will not just contain
-code pointers and data.  This means that this discipline should be combined with
-protection against stack smashing attacks (i.e. buffer overflows on the stack overwriting activation
-records' contents), but capability machines' fine-grained memory protection
-luckily is well-suited for this.
-
-\textbf{Invoking an adversary callback}
-* must be global
+code pointers and data. This means that our discipline should be combined with
+protection against stack smashing attacks (i.e. buffer overflows on the stack
+overwriting activation records' contents), but conveniently, the fine-grained
+memory protection on capability machines is well-suited to easily provide such
+protection.
+
+\textbf{Invoking an adversary callback} If we have a higher-order interface to
+the adversary, we might need to invoke an adversary callback. In this case, not
+so much changes with respect to the situation where we invoke regular adversary
+code. The adversary can provide a callback as a capability for us to jump to,
+either an $\entry$-capability if he wants to protect himself from us or just an
+$\exec$ capability if he/she is not worried about that. However, there is one
+scenario that we need to prevent, namely that he/she constructs the callback
+capability to point into the stack, where he/she may have stored local
+capabilities that he should not have access to upon invocation of the callback.
+Specifically, he might store return pointers and stack pointers from previous
+stack frames there, and illegally use them inside the callback.
+
+To prevent this, we only accept callbacks from the adversary in the form of
+global capabilities and we check this prior to invoking them (and fail
+otherwise). This should not be an overly strict requirement, since our own
+callbacks do not contain local data themselves, so there should be no need for
+the adversary to construct callbacks on the stack.\footnote{Note that it does
+  prevent a legitimate but non-essential scenario where the adversary wants to
+  give us temporary access to a callback not allocated on the stack.}
 
 \paragraph{Having a callback invoked by the adversary}
-* checking the received stack pointer
-
+The above leaves us with probably the hardest scenario: how to provide a
+callback to the adversary. The basic idea is that we allocate a block of memory
+using malloc that we fill with the capabilities and data that the callback needs
+access to, as well as some prelude instructions that load the data into
+registers and jumps to the right code. We can then provide the adversary with an
+enter-capability covering the allocated block and pointing to the contained
+prelude instructions. However, the question that remains in this setup is: where
+do we get the stack pointer?
+
+Our answer is that the adversary should provide it to us, much like we provide
+him with a current stack pointer when we invoke his code.  However, it is
+important that we do not just accept any capability as a stack pointer but check
+that it is safe to use.  Specifically, we check that it is indeed a $\rwlx$
+pointer and that its current address points just below its start address.
+Without these checks, err...
+\dominique{is this needed for security or just for convenience?}
+
+TODO:
 \begin{itemize}
-\item Issues to consider:
-\begin{itemize}
-\item no global store-local capabilities
-\item clearing the stack
-\item clearing (some) registers
-\item constructing return pointers
-\item constructing callbacks
-\item invoking the adversary's callbacks
-\end{itemize}
 \item Reusable macro-instructions for stack and return pointer management

### 2017-02-10 csf-2017/paper.tex
more about examples

--- 2017-02-10 csf-2017/paper.tex
@@
 % One argument
-\newcommand{\oneinstr}[2]{
-  \ifthenelse{\equal{#2}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2}
-}
+\newcommand{\oneinstr}[2]{\zinstr{#1} \; #2}
 \newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
 % Two arguments
-\newcommand{\twoinstr}[3]{
-  \ifthenelse{\equal{#2#3}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3}
-}
+\newcommand{\twoinstr}[3]{\zinstr{#1} \; #2 \; #3}
 \newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
@@
 % Three arguments
-\newcommand{\threeinstr}[4]{
-  \ifthenelse{\equal{#2#3#4}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3 \; #4}
-}
+\newcommand{\threeinstr}[4]{\zinstr{#1} \; #2 \; #3 \; #4}
 \newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
@@
 three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
-address), $\restricttwo{}{}$ (modifies the permission and local/global flag) and
+address), $\restrict{}{}$ (modifies the permission and local/global flag) and
 $\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
@@
 into the appropriate register, but only if the capability has appropriate
-authority (i.e. read permission and an appropriate range). The $\restricttwo{}{}$
+authority (i.e. read permission and an appropriate range). The $\restrict{}{}$
 instruction will update a capability's permissions and global/local status in
@@
 
-Explain malloc
-\dominique{not sure what we want to say about it here either...}
-
 \section{Stack and return pointer management using local capabilities}
@@
 
-The basic idea of our approach is quite simple: we stick to a single, rather
-standard C stack and register-passed return pointers in much the same way as a
-standard C ABI. However, to prevent various ways of misusing this basic scheme,
-we put local capabilities to work and take a number of not-always-obvious safety
-measures. In the next paragraphs, we will explain the issues to be considered in
-all the relevant situations: when (1) starting our program, (2) invoking the
-adversary, (3) returning from the adversary, (4) invoking an adversary callback
-and (5) having a callback invoked by the adversary.
-
-\dominique{Add a summary table somewhere?}
-
-\paragraph{Program startup}
-When our code starts executing, we assume that the language runtime initializes
-the memory as follows: a contiguous array of memory is reserved for the stack,
-for which we receive a stack pointer in a register that we reserve for this
-purpose \dominique{give this register a special name like sp?}. This stack
-pointer must be local and it must carry $\rwlx$ permission. It is crucial for
-our approach that the stack is the only part of memory for which there will ever
-be capabilities with $\rwlx$ or $\rwl$ capabilities. We expect the language
-runtime to enforce this by not handing out such capabilities for other parts of
-memory in any situation (i.e. not through malloc, not during loading or linking
-etc.). Additionally, in our examples, we typically also assume some static parts
-of memory where instructions are stored or static data. Another part of memory
-(called the heap) is initially governed by malloc and at program startup, no
-other code has capabilities for this memory. Malloc will hand out capabilities
-for allocated regions as requested, but it will never hand out capabilities with
-write-local capability (i.e. no $\rwlx$ or $\rwl$ permissions). For simplicity,
-we assume that memory allocated through malloc cannot be freed.
+The basic idea of our approach is actually rather simple: we stick to a single,
+rather standard C stack and register-passed return pointers in much the same way
+as a standard C ABI. However, to prevent various ways of misusing this basic
+scheme, we put local capabilities to work and need to perform a number of
+not-always-obvious checks and safety measures.
 
 \paragraph{Invoking the adversary}
-During execution, we use the stack as normal: we may push values onto it and pop
-them back off (e.g. for register spilling etc.). When we invoke the adversary,
-we need to take great care to only provide him with the authority we want him to
-have. Specifically, we clear all the registers, except those that contain
-arguments the adversary is intended to have access to. We leave a copy of the
-stack pointer in the appropriate register, but only after we have used the
-$\subseg{}{}{}$ to shrink its authority to just the part of the stack that we
-are not using. In one of the registers, we will also provide a return pointer.
-We will go into constructing the return pointer further below.
-
-It is already important to note that the return pointer
-will be provided to the adversary as local capabilities. If the adversary were
-able to store away the return pointer in a global data structure, then he would
-be able to jump to it in circumstances where this should not be possible: he
-may, for example, store the return pointer, legally jump to it a first time,
-wait to be invoked again and then jump to the old return pointer a second time,
-rather than the new return pointer received for the new invocation. Similarly,
-he may store the return pointer, invoke a function in the trusted code, wait to
-be invoked again and then jump to the stored return pointer rather than the new
-return pointer for the second invocation.
-
-Because the return pointer is local, the adversary is only able to store it
-through write-local capabilities, which (because of our assumptions above)
-means: on the stack. However, this makes it crucial that the stack pointer is
-also local: if the adversary were able to store the stack pointer in a global
-data structure, he/she can store return pointers on the stack, and play similar
-mischief as mentioned above. Storing away stack pointers for use during future
-invocations would also be dangerous for other reasons: during such a future
-invocation, a larger part of the stack may be in use by the trusted code, which
-the adversary may still have access to through the old stack pointer.
-
-A final measure we need to take is that before invocation of the adversary, we
-need to clear (i.e. overwrite with zeros) the part of the stack that we are not
-using.  Otherwise, that part of the stack may still contain data and
-capabilities from previously executed code, including local capabilities since
-they are allowed on the stack.  The adversary would be able to take advantage of
-this by using it to store either stack pointers or return pointers from previous
-invocations and play the same tricks as above.
+* stack and return pointer as local caps
+* clearing registers
+* clearing the stack
 
 \paragraph{Returning from the adversary}
-So return pointers must be passed as local capabilities. But what should their
-permissions be, what memory should they point to and what should that memory
-(let's call it the activation record) contain? Let us answer the last question
-first by considering what should happen when the adversary jumps to a return
-pointer and what capabilities we need for achieving that. Essentially, the
-program counter should be restored to the code after the call into the
-adversary, so the activation record should store this program counter.
-Additionally, the stack pointer should also be restored to its original value.
-Since the adversary has a more restricted authority over the stack than the code
-making the call, we cannot hope to reconstruct the return pointer from the stack
-pointer owned by the adversary. Instead, the old stack pointer should also be
-stored as part of the activation record.
-
-\begin{figure}
-  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
-    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
-    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
-    \draw (0,0.5) -- (0,4);
-    \draw (6,0.5) -- (6,4);
-    \draw (0,3.5) rectangle (6,3) node[pos=.5] {\footnotesize previous program counter};
-    \draw (0,3) rectangle (6,2.5) node[pos=.5] {\footnotesize previous stack pointer};
-    \draw (0,2.5) rectangle (6,1) node[pos=.5] {\footnotesize restore instructions};
-    \draw (-3,2.25) node (rp) {return pointer};
-    \draw[->] (rp.east) -- (-.2,2.25);
-  \end{tikzpicture}
-
-  \caption{Structure of an activation record}
-  \label{fig:activ-rec-struct}
-\end{figure}
-
-It's important that neither of these capabilities are accessible to the
-adversary. In other words, we need to provide the adversary a capability that he
-can jump to but not read from, i.e. an enter capability. To make this work, we
-construct the activation record as depicted in Figure~\ref{fig:activ-rec-struct}.
-
-* return pointer as enter pointer to activation
-record on the stack. * activation record contains previous pc and stack pointer
+* return pointer as enter pointer to activation record on the stack.
+* activation record contains previous pc and stack pointer
 
@@
 explained when needed. The precise definitions can be found in the
-technical report~\cite{XXX}.
+technical appendix~\cite{XXX}.
 
@@
 local state encapsulation and control-flow integrity properties. For
-presentation purposes, we do not formally state the lemmas we
-prove. 
-\lars{I think it would be good if we can state one of them,
-  possibly in shortened form, just to make it clear that something is
-  actually stated and proved formally, and make a link to the LR}
+presentation purposes, we omit some details in the lemmas we state. 
 The interested reader can find all the technical details in the
-techincal report.
+techincal report~\cite{XXX}.
 
@@
 (described in the technical report) to run the adveserial code.
+
+\begin{figure}[htbp]
+  \centering
+  \begin{lstlisting}
+f1: push 1
+    fetch $r_1$ $\adv$
+    scall $r_1$($[]$,$[]$)
+    pop $r_1$
+    assert $r_1$ 1
+    halt
+\end{lstlisting}
+  \caption{Simple program that uses a stack-based calling convention.}
+  \label{fig:prog-f1}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+\begin{lstlisting}
+f2: malloc $r_l$ 1
+    store $r_l$ 1
+    fetch $r_1$ $\adv$
+    call $r_1$($[]$,$[r_l]$)
+    assert $r_l$ 1
+    halt
+\end{lstlisting}
+  \caption{Simple program that uses a calling convention not relyant
+    on the stack}
+  \label{fig:prog-f2}
+\end{figure}
 
@@
   \[
-    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[*] (\halted,m'),
+    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m'),
   \]
@@
 
-For \texttt{f2} we state a similar lemma:
+We have a similar lemma for \texttt{f2}:
 \begin{lemma}
@@
   Lemma~\ref{lem:correctness-f1} but assuming no stack capability
-  if $(\reg\update{\pcreg}{c_{f2}},m) \step[n] (\halted,m')$, then $m'(\flag) = 0$.
+  if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
 \end{lemma}
-
-\begin{figure}[htbp]
-  \centering
-  \begin{lstlisting}
-f1: push 1
-    fetch $r_1$ $\adv$
-    scall $r_1$($[]$,$[]$)
-    pop $r_1$
-    assert $r_1$ 1
-    halt
-\end{lstlisting}
-  \caption{Simple program that uses a stack-based calling convention.}
-  \label{fig:prog-f1}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-\begin{lstlisting}
-f2: malloc $r_l$ 1
-    store $r_l$ 1
-    fetch $r_1$ $\adv$
-    call $r_1$($[]$,$[r_l]$)
-    assert $r_l$ 1
-    halt
-\end{lstlisting}
-  \caption{Simple program that uses a calling convention not relyant
-    on the stack}
-  \label{fig:prog-f2}
-\end{figure}
 
@@
 
-The final example in \figurename~\ref{fig:prog-g1} consists of two
-part: \texttt{g1} is a closure generation program with one variable $x$ in its
-environment and \texttt{f4} as the code.
-Like in the previous example, the assertion can fail if the adversary can capture the protected return pointer of the second call, invoke the closure anew and use the captured protected return pointer in the first call when the variable $x$ has the value 0.
-
-In the other examples, the known program started the execution and called an unknown piece of code.
-In this example, the control is inverted, so an unknown piece of code gets an enter capability for the closure creator.
-
-For the property we show for \texttt{g1}, the control is inverted
-compared to the other examples.
+\begin{lemma}
+  \label{lem:correctness-f3}
+  Making similar assumptions about capabilities and linking as in
+  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
+  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
+  (\halted,m')$, then $m'(\flag) = 0$.
+\end{lemma}
+
+The final example, in \figurename~\ref{fig:prog-g1}, is a faithful
+translation of a trciky example known from the
+litterature\cite{XXX}. It consists of two parts, \texttt{g1} and
+\texttt{f4}. \texttt{g1} is a closure generator that generates a
+closure with one variable $x$ set to 0 in its environment and \texttt{f4} as
+the program. The program \texttt{f4} expects to arguments: a stack and
+a callback. The stack is necessary to use \texttt{scall}, but because
+it comes from an untrusted source, we need to make sure that it has
+the correct permission and fits with our stack dicipline. The callback
+must be global for the program to work, we will get back to this
+point. The program now sets $x$ to 0 and calls the callback. When it
+returns from the callback, it sets $x$ to 1 and calls the callback a
+second time. When it returns from the callback, it asserts $x$ is 1
+and returns.
+
+To illustrate how subtle this program is, consider one way an
+adversary may try to make the assertion fail. In the second callback
+an adversary can get to the first callback by calling the closure one
+more time. If the adversary can save the protected return pointer in
+the second callback and use it in the first, then the assertion will
+fail because $x$ is 0 before the first callback. The protected return
+pointer cannot be transfered between the callbacks, so this does not
+work.
+
+\begin{lemma}
+  \label{lem:correctness-g1}
+  Making similar assumptions about capabilities and linking as in
+  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
+  if
+  $(\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m')$,
+  then 
+\[
+  m'(\flag) = 0.
+\]
+\end{lemma}
+
+As mentioned, the callback must be global. This is to make sure the
+callback is not a piece of code on the stack that restores the
+adversaries original stack capability breaking the encapsulation of
+our local stack frame. In the proof of Lemma~\ref{lem:correctness-g1},
+this also shows up. When we reason about \texttt{f4}, we revoke all
+the local state of the adversary which means that we consider a
+private future world. We use the $\execCond{}$ to reason about the
+jump to the callback and because the callback is global, it works for
+private future worlds. Had the callback been local, then the
+$\execCond{}$ would only work for public future worlds, and we would
+be stuck in the proof.
+
+In Lemma~\ref{lem:correctness-g1} the order of control has been
+inverted compared to the previous lemmas. In this lemma, the adversary
+assumes control first with a capability for the closure creator
+\texttt{g1}. This means we need to check all arguments are safe
+to use and that we clean up before returning in the end. The inversion
+of control poses an interesting challenge when it comes to reasoning
+bout the adversaries local state during the execution of
+\texttt{f4}. During the execution of \texttt{f4} and the callbacks the
+adversary should not be able to rely on its local state from outside
+the call of \texttt{f4}. This is easily done by revoking all the
+temporary regions of the world given at the start of \texttt{f4}. The
+real challenge comes when \texttt{f4} returns and needs to guarantee
+that the protocols of the adversaries local state has not been
+violated. To this end, we need to reinstate regions we revoked which
+is done using the masking regions.
 

### 2017-02-10 csf-2017/paper.tex
he/she to they

@@
 for the duration of an invocation: if we take care to erase the capability from
-the register file after control is passed back to us, we can be sure that he/she
+the register file after control is passed back to us, we can be sure that they
 will not have been able to store the capability. However, there is one exception
@@
 \textbf{Returning to the adversary} Perhaps the simplest scenario is the case of
-returning to the adversary after he/she invoked our code. In this case, we will
+returning to the adversary after they invoked our code. In this case, we will
 have received a return pointer from him, and we just need to jump to it as
@@
 stack pointer is also local: if the adversary were able to store the stack
-pointer in a global data structure, he/she can store return pointers on the
+pointer in a global data structure, they can store return pointers on the
 stack, and play similar mischief as mentioned above. Storing away stack pointers
@@
 adversary. In other words, the return pointer provided to the adversary must be
-a capability that he/she can jump to but not read from, i.e. an enter
+a capability that they can jump to but not read from, i.e. an enter
 capability. To make this work, we construct the activation record as depicted in
@@
 register file. Because the return pointer is given to the adversary as an enter
-pointer, he/she cannot get hold of the contents of the data structure, but after
+pointer, they cannot get hold of the contents of the data structure, but after
 invocation, its permission is updated to $\exec$, so the activation record's
@@
 either an $\entry$-capability if he wants to protect himself from us or just an
-$\exec$ capability if he/she is not worried about that. However, there is one
-scenario that we need to prevent, namely that he/she constructs the callback
-capability to point into the stack, where he/she may have stored local
+$\exec$ capability if they is not worried about that. However, there is one
+scenario that we need to prevent, namely that they constructs the callback
+capability to point into the stack, where they may have stored local
 capabilities that he should not have access to upon invocation of the callback.

### 2017-02-10 csf-2017/paper.tex
pass over examples

@@
 \section{Examples}
-In this section, we display how capability safety allows us to prove
-local state encapsulation and control-flow integrity properties. For
-presentation purposes, we omit some details in the lemmas we state. 
-The interested reader can find all the technical details in the
-techincal report~\cite{XXX}.
+In this section, we demonstrate how our formal model of capability
+safety allows us to prove local state encapsulation and control-flow
+integrity properties for challenging examples. 
+Since we are dealing with assembly language, 
+there are many details to the formal treatment , and therefore we 
+necessarily have to omit some details
+in the lemmas we state here.  The interested reader can find all the
+technical details in the techincal report~\cite{XXX}.
 
@@
 The capability machine provides encapsulation of local state. The
-program \texttt{f1} in \figurename~\ref{fig:prog-f1} and the program
+programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and 
 \texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.
-
-The programs \texttt{f1} and \texttt{f2} are very similar. They both
-store some local data, call an untrusted piece of code, and finally
+The two programs \texttt{f1} and \texttt{f2} are very similar. They both
+store some local data, call an untrusted piece of code ($\adv$), and finally
 make sure that their local state is unchanged. They differ in the way
-they do this. \texttt{f1} assumes it has access to a stack capability
+they do this. Program \texttt{f1} assumes it has access to a stack capability
 and uses the stack to store its local state. It also uses the
@@
 state and uses an activation-record based calling convention
-(described in the technical report) to run the adveserial code.
+(described in the technical report) to run the adversarial code.
 
@@
 
-For both programs, we can prove that if they are linked up with an
+For both programs, we can prove that if they are linked with an
 adversary, $\adv$, that is allowed to allocate new memory but otherwise has no
-capabilities, then executing the program, the assertion will never
-fail.
+capabilities, then the assertion will never fail when
+executing the program,
+
 We can only prove this, because the capability machine encapsulates
@@
 
-\todo[inline]{give a more detailed high-level taste of a proof here? Maybe later for one of the other examples?}
-\begin{lemma}[Correctness lemma for \texttt{f1}]
-\todo[inline]{The lemma copied directly from the TR. To be deleted later!}
-  \label{lem:correctness-f1}
-  Let
-  \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
-    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
-    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
-    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
-    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
-    \reg & \in \Regs \\
-    m & \defeq \hs_{f1} \uplus 
-        \hs_\flag \uplus                
-        \ms_{\var{link}} \uplus 
-        \hs_\adv \uplus 
-        \ms_{\malloc} \uplus 
-        \ms_{\var{stk}} \uplus
-        \ms_{\var{frame}} 
-  \end{align*}
-  and
-  \begin{itemize}
-  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
-  \end{itemize}
-  where 
-  \begin{align*}
-    &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
-    &\dom(\hs_\flag) = [\flag,\flag] \\
-    &\dom(\ms_\link) = [\link,\link+1]\\
-    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
-    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
-  \end{align*}
-  and
-  \begin{itemize}
-  \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
-  \item $\ms_\flag = [\flag \mapsto 0]$
-  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-  \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
-  \end{itemize}
-  if 
-  \[
-    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
-  \]
-  then
-  \[
-    m'(\flag) = 0
-  \]  
-\end{lemma}
+% \todo[inline]{give a more detailed high-level taste of a proof here? Maybe later for one of the other examples?}
+% \begin{lemma}[Correctness lemma for \texttt{f1}]
+% \todo[inline]{The lemma copied directly from the TR. To be deleted later!}
+%   \label{lem:correctness-f1}
+%   Let
+%   \begin{align*}
+%     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+%     c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
+%     c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+%     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+%     c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
+%     \reg & \in \Regs \\
+%     m & \defeq \hs_{f1} \uplus 
+%         \hs_\flag \uplus                
+%         \ms_{\var{link}} \uplus 
+%         \hs_\adv \uplus 
+%         \ms_{\malloc} \uplus 
+%         \ms_{\var{stk}} \uplus
+%         \ms_{\var{frame}} 
+%   \end{align*}
+%   and
+%   \begin{itemize}
+%   \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
+%   \end{itemize}
+%   where 
+%   \begin{align*}
+%     &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
+%     &\dom(\hs_\flag) = [\flag,\flag] \\
+%     &\dom(\ms_\link) = [\link,\link+1]\\
+%     &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+%     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+%     &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
+%   \end{align*}
+%   and
+%   \begin{itemize}
+%   \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
+%   \item $\ms_\flag = [\flag \mapsto 0]$
+%   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+%   \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
+%   \end{itemize}
+%   if 
+%   \[
+%     (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+%   \]
+%   then
+%   \[
+%     m'(\flag) = 0
+%   \]  
+% \end{lemma}
 

### 2017-02-10 csf-2017/paper.tex
added minus instruction

--- 2017-02-10 csf-2017/paper.tex
@@
 \newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}
+\newcommand{\minus}[3]{\threeinstr{minus}{#1}{#2}{#3}}
 
@@
   novel technical characteristics. The natural CPS style of assembly
-  code with return pointers as continuations is modelled using a
+  code with return pointers as continuations is modeled using a
   \emph{single} orthogonal closure, rather than the biorthogonal
@@
            &   & \plus{\lv}{\rv}{\rv} \mid 
+                 \minus{\lv}{\rv}{\rv} \mid 
                  \lea{\lv}{\rv} \mid 
@@
 load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and an arithmetic addition operator ($\plus{}{}{}$). There are
+and $\store{}{}$) and an arithmetic addition operators ($\plus{}{}{}$ and $\minus{}{}{}$). There are
 three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
@@
 \end{figure}
-A world is a finite map from region names, modelled as natural numbers, to regions.
+A world is a finite map from region names, modeled as natural numbers, to regions.
 We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.

### 2017-02-10 csf-2017/paper.tex
pass over examples

--- 2017-02-10 csf-2017/paper.tex
@@
 The two programs \texttt{f1} and \texttt{f2} are very similar. They both
-store some local data, call an untrusted piece of code ($\adv$), and finally
-make sure that their local state is unchanged. They differ in the way
+store some local data, call an untrusted piece of code ($\adv$), and
+then tests whether local state is unchanged. They differ in the way
 they do this. Program \texttt{f1} assumes it has access to a stack capability
@@
 capabilities, then the assertion will never fail when
-executing the program,
-
-We can only prove this, because the capability machine encapsulates
-the local state of the two programs. The two examples also illustrate
+executing the program (see Lemmas~\ref{lem:correctness-f1}
+and~\ref{lem:correctness-f2} below).
+\lars{shouldn't we say already here that the adversary is a global 
+  capability and that is why, intuitively, the local state is
+  preserved ?}
+The two examples also illustrate
 the versatility of the logical relation. The two examples use
@@
 
-% \todo[inline]{give a more detailed high-level taste of a proof here? Maybe later for one of the other examples?}
 % \begin{lemma}[Correctness lemma for \texttt{f1}]
@@
 
-In order to formulate a lemmas about \texttt{f1} and \texttt{f2}, we need a way to observe whether the assertion fails.
-To this end, we assum access to a flag which will be set by the assertion if it fails.
-A slightly simplified version of the correctness lemma for \texttt{f1} is:
+In order to formulate lemmas about \texttt{f1} and \texttt{f2}, we
+need a way to observe whether the assertion fails. To this end, we
+assume access to a flag (an address in memory)
+which will be set to $1$ by the assertion if it
+fails. A slightly simplified version of the correctness lemma for
+\texttt{f1} is:
 \begin{lemma}
@@
 
-In order for prove Lemma~\ref{lem:correctness-f1}, it suffices to show
+To prove Lemma~\ref{lem:correctness-f1}, it suffices to show
 that the start configuration is capability safe (in the $\observations$ relation) for a
@@
 show that it is in it at a later point of the execution. To reason
-about \texttt{scall}, it suffices to show that the configuration just
-after we jumped is capability safe and that the configuration just
+about \texttt{scall}, it suffices to show that (1) the configuration just
+after we jumped is capability safe and (2) that the configuration just
 after \texttt{scall} is done cleaning up is capability safe.

### 2017-02-10 csf-2017/paper.tex
changed how we write registers for instructions

--- 2017-02-10 csf-2017/paper.tex
@@
 \usepackage{rotating}
-
 \usepackage{afterpage}
@@
 \newcommand{\ms}{\hs}
-\newcommand{\hv}{\var{hv}}
-\newcommand{\rv}{\var{rv}}
-\newcommand{\lv}{\var{lv}}
+\newcommand{\hv}{\var{r}}
+\newcommand{\rv}{\var{r}}
+\newcommand{\lv}{\var{r}}
 \newcommand{\gl}{\var{g}}
@@
 % Reference register/memory
-\newcommand{\refreg}[1]{\lfloor #1 \rfloor}
-\newcommand{\refheap}[1]{\langle #1 \rangle_m}
+\newcommand{\refreg}[1]{#1}
+\newcommand{\refheap}[1]{#1}
 
@@
   \begin{array}{rcl}
-    \lv    &::=& \refreg{r} \\
-    \hv    &::=& \refheap{r}\\
     \rv    &::=& n \mid \lv \\
@@
                  \lea{\lv}{\rv} \mid 
-                 \restricttwo{\lv}{\rv} \mid 
-                 \subseg{\lv}{\rv}{\rv} \mid  \\
-           &   & \isptr{\lv}{\rv} \mid 
+                 \restricttwo{\lv}{\rv} \mid \\
+           &   & \subseg{\lv}{\rv}{\rv} \mid  
+                 \isptr{\lv}{\rv} \mid 
                  \getp{\lv}{\lv} \mid 
-                 \getl{\lv}{\lv} \mid 
-                 \getb{\lv}{\lv} \mid
-                 \gete{\lv}{\lv} \mid\\
-           &   & \geta{\lv}{\lv} \mid 
+                 \getl{\lv}{\lv} \mid\\ 
+           &   & \getb{\lv}{\lv} \mid
+                 \gete{\lv}{\lv} \mid
+                 \geta{\lv}{\lv} \mid 
                  \fail \mid
@@
     \begin{cases}
-      \sem{\decode(n)}(\Phi) & \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }\\
-      \failed                                 & \text{otherwise}
+      \sem{\decode(n)}(\Phi) & \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }}\\
+      \failed                                 & \text{\footnotesize{otherwise}}
     \end{cases}
@@
                        \begin{cases}
-                         ((\exec,\gl),\start,\addrend,\addr) & \text{if $w = ((\entry,\gl),\start,\addrend,\addr)$}\\
-                         w & \text{otherwise}
+                         ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
+                         w & \text{\footnotesize{otherwise}}
                        \end{cases}\\
@@
                            \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
-                           \begin{array}[t]{l}
-                             \text{if $\memreg(\pcreg) = \stdcap$}\\
-                             \quad\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}\\
-                           \end{array}
+                             \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}\\
+                             \failed & \text{\footnotesize{otherwise}}
                          \end{cases}
@@
     \hline
-    $\move{\refreg{r_1}}{\rv}$& $\begin{multlined}
+    $\move{\refreg{r_1}}{r_2}$& $\begin{multlined}
       \stdUpdatePc{}(\updateReg{r_1}{\\
-          \memreg(\rv)}) \end{multlined}$&\\
+          \memreg(r_2)}) \end{multlined}$&\\
     \hline
-    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$  and  $\start \leq \addr \leq \addrend$\\
+    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$\footnotesize{ and }$\start \leq \addr \leq \addrend$\\
     \hline
-    $\restricttwo{\refreg{r}}{\rv}$&$\begin{multlined}
-                                       \stdUpdatePc{}(\updateReg{r}{\\(\permp',\start,\addrend,\addr)})
+    $\restricttwo{\refreg{r_1}}{r_2}$&$\begin{multlined}
+                                       \stdUpdatePc{}(\updateReg{r_1}{\\(\permp',\start,\addrend,\addr)})
                                      \end{multlined}$  &
-                                                        $\memreg(r) = \stdcap[\permp]$ and $\permp' = \decodePermPair{\memreg(\rv)}$
-                                                      and $\permp'\sqsubseteq \permp$\\
+                                                        $\memreg(r_2) = \stdcap[\permp]$\footnotesize{ and }\newline$\permp' = \decodePermPair{\memreg(r_2)}$
+                                                      \footnotesize{ and }$\permp'\sqsubseteq \permp$\\
     \hline
@@
     \hline
-    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and  $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
-                                                                                       and if  $\var{w} = ((\_,\local),\_,\_,\_)$, then  $\perm \in \{\rwlx,\readwritel \}$\\
+    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$ \footnotesize{ and } $\start \leq \addr \leq \addrend$\footnotesize{ and }$\var{w} = \memreg(r_2)$
+                                                                                       \footnotesize{and if }$\var{w} = ((\_,\local),\_,\_,\_)$\footnotesize{, then  }$\perm \in \{\rwlx,\readwritel \}$\\
     \hline
@@
     \hline
-    \_&$\failed$&otherwise\\
+    \_&$\failed$&\footnotesize{otherwise}\\
     \hline

### 2017-02-10 csf-2017/paper.tex

--- 2017-02-10 csf-2017/paper.tex
@@
            &   & \plus{\lv}{\rv}{\rv} \mid 
+                 \minus{\lv}{\rv}{\rv} \mid 
                  \lea{\lv}{\rv} \mid 
-                 \restricttwo{\lv}{\rv} \mid 
-                 \subseg{\lv}{\rv}{\rv} \mid  \\
-           &   & \isptr{\lv}{\rv} \mid 
+                 \restricttwo{\lv}{\rv} \mid \\
+           &   & \subseg{\lv}{\rv}{\rv} \mid  
+                 \isptr{\lv}{\rv} \mid 
                  \getp{\lv}{\lv} \mid 
-                 \getl{\lv}{\lv} \mid 
-                 \getb{\lv}{\lv} \mid
-                 \gete{\lv}{\lv} \mid\\
-           &   & \geta{\lv}{\lv} \mid 
+                 \getl{\lv}{\lv} \mid\\ 
+           &   & \getb{\lv}{\lv} \mid
+                 \gete{\lv}{\lv} \mid
+                 \geta{\lv}{\lv} \mid 
                  \fail \mid

### 2017-02-10 csf-2017/paper.tex
first comments and changes to capability machine section (and a few to the one after)

@@
 
-We assume a set of register names $\RegName$. We define register files $\reg$ as
+We assume a finite set of register names $\RegName$. We define register files $\reg$ as
 functions mapping register names to words and memories as functions that map
@@
 parts of memory represented as partial functions mapping addresses to words.
+\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
 
@@
 
-Figure~\ref{fig:op-sem} shows the operational semantics of our capability
-machine, or at least an excerpt containing a number of representative
+\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics containing a number of representative
 instructions. Essentially, a configuration $\Phi$ will either decode and execute
@@
 
-The $\jmp{}$ instruction updates the program counter to a requested location,
-but it is complicated by the presence of \emph{enter capabilities}, modeled
-after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
-cannot be used to read, write or execute and their current address and range
-cannot be modified. They can only be used to jump to, but when that happens,
-their permission is updated to $\exec$. This special behavior makes them
-suitable for representing a kind of closures: an opaque package containing a
-piece of code together with local encapsulated state. The idea is that one can
-build an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where the
-range $[\start,\addr-1]$ is used for storing the closure's local state (data or
-capabilities) and the range $[\addr,\addrend]$ can be used for storing
-instructions. The instructions and data are opaque to an adversary holding $c$
-but when $c$ jumped to, the data becomes available to the closure's code through
-the updated version of $c$ that is now in the $\pcreg$\footnote{Note that
-  $\pcreg$ can be used to $\load{}{}$ from, like any other register.}.
+The $\jmp{}$ instruction updates the program counter to a requested
+location, but it is complicated by the presence of \emph{enter
+  capabilities}, modeled after the
+M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
+cannot be used to read, write or execute and their current address and
+range cannot be modified. They can only be used to jump to, but when
+that happens, their permission is updated to $\exec$. This special
+behavior makes them suitable for representing a kind of closures: an
+opaque package containing a piece of code together with local
+encapsulated state. The idea is that one can build an enter capability
+$c = ((\entry,\gl),\start,\addrend,\addr)$ where the range
+$[\start,\addr-1]$ is used for storing the closure's local state (data
+or capabilities) and the range $[\addr,\addrend]$ can be used for
+storing instructions. The instructions and data are opaque to an
+adversary holding $c$ but when $c$ jumped to, the data becomes
+available to the closure's code through the updated version of $c$
+that is now in the $\pcreg$\footnote{Note that the capability in the
+  $\pcreg$-register can be moved to another register and used to load
+  from.}. \lau{As we talked about what is described here, is not how we make closures in the TR, so we should consider whether we should refrain from using that word here. }
 
@@
 \section{Stack and return pointer management using local capabilities}
-
+\lau{Do we want to use the term ``pointer'' at any point? I have also used it, but I guess it might create some confusing to whether we have traditional pointer as well as capabilities?}
 One of the contributions in this paper is a demonstration that local
@@
 the memory as follows: a contiguous array of memory is reserved for the stack,
-for which we receive a stack pointer in a register that we reserve for this
-purpose \dominique{give this register a special name like sp?}. This stack
+for which we receive a stack pointer in a register, $r_\stk$, that we reserve for this
+purpose. This stack
 pointer must be local and it must carry $\rwlx$ permission. It is crucial for
 our approach that the stack is the only part of memory for which there will ever
-be capabilities with $\rwlx$ or $\rwl$ capabilities. We expect the language
+be an obtainable capabilities with $\rwlx$ or $\rwl$ capabilities. We expect the language
 runtime to enforce this by not handing out such capabilities for other parts of

### 2017-02-10 csf-2017/paper.tex
added comments for macr section

@@
 standard C stack and register-passed return pointers in much the same way as a
-standard C ABI. However, to prevent various ways of misusing this basic scheme,
+standard C ABI\lau{This sentence does not compute}. However, to prevent various ways of misusing this basic scheme,
 we put local capabilities to work and take a number of not-always-obvious safety
@@
 returning to the adversary after they invoked our code. In this case, we will
-have received a return pointer from him, and we just need to jump to it as
+have received a return pointer from them, and we just need to jump to it as
 usual. The only thing we need to take care of is to properly clear the registers
@@
 away the return pointer in a global data structure, and he would be able to jump
-to it in circumstances where this should not be possible: he may, for example,
+to it in circumstances where this should not be possible: for example, he may
 store the return pointer, legally jump to it a first time, wait to be invoked
@@
 and then jump to the stored, old return pointer rather than the new one he
-received upon the second invocation.
+received upon the second invocation. \lau{If we need space, then maybe one of the examples is sufficient.}
 
@@
 CHERI should consider to provide special support for this requirement, by
-implementing an optimized instruction for clearing (i.e. overwriting with zeros)
+implementing an optimized instruction for clearing
 a large block of memory. We have no experience in processor design, but we
@@
 
-\paragraph{Having a callback invoked by the adversary}
+\textbf{Having a callback invoked by the adversary}
 The above leaves us with probably the hardest scenario: how to provide a
@@
 \dominique{is this needed for security or just for convenience?}
+\lau{Say we are called with a stack capability that is global and $\rwx$ as well as an integer as a non-local return pointer (this is fair game as the adversary has to follow no rules).
+Now you push the l(a)unch codes to what you think is the stack and \texttt{scall} the callback.
+At this point \texttt{scall} does its thing, but because it never attempts to store anything local it will succeed and jump to the adversary!
+The adversary can now look at what you thought was the stack and eat your lunch or fire the missiles as they prefer.
+It is not sufficient to require the stack to be just local because there may be a global equivalent.
+We know this is not the case for the rwlx capability because we have a system wide assumption that all rwlx capabilities are local.
+Finally the adjustment of the pointer in \texttt{prepstck} (or whatever I called it) is a convenience/ space efficiency thing purely.}
 
@@
 
-LB: maybe we can present the examples informally before the logical relation and then return to them after the logical relation (to present the formal properties) ?
+\lars{maybe we can present the examples informally before the logical relation and then return to them after the logical relation (to present the formal properties) ?}
 

### 2017-02-10 csf-2017/paper.tex
grammar

@@
 adversary as a local capability. If not, the adversary would be able to store
-away the return pointer in a global data structure, and he would be able to jump
-to it in circumstances where this should not be possible: for example, he may
+away the return pointer in a global data structure, and they would be able to jump
+to it in circumstances where this should not be possible: for example, they may
 store the return pointer, legally jump to it a first time, wait to be invoked
 again and then jump to the old return pointer a second time, rather than the new
-return pointer received for the new invocation. Similarly, he may store the
+return pointer received for the new invocation. Similarly, they may store the
 return pointer, invoke a function in our code, wait for us to invoke him again
@@
 code. The adversary can provide a callback as a capability for us to jump to,
-either an $\entry$-capability if he wants to protect himself from us or just an
-$\exec$ capability if they is not worried about that. However, there is one
-scenario that we need to prevent, namely that they constructs the callback
+either an $\entry$-capability if they want to protect themselves from us or just an
+$\exec$ capability if they are not worried about that. However, there is one
+scenario that we need to prevent, namely that they construct the callback
 capability to point into the stack, where they may have stored local
-capabilities that he should not have access to upon invocation of the callback.
-Specifically, he might store return pointers and stack pointers from previous
+capabilities that they should not have access to upon invocation of the callback.
+Specifically, they might store return pointers and stack pointers from previous
 stack frames there, and illegally use them inside the callback.
@@
 Our answer is that the adversary should provide it to us, much like we provide
-him with a current stack pointer when we invoke his code.  However, it is
+them with a current stack pointer when we invoke his code.  However, it is
 important that we do not just accept any capability as a stack pointer but check

### 2017-02-10 csf-2017/paper.tex
Optimize math layout a bit, remove obsolete remark.

@@
 \newcommand{\npair}[2][n]{\left(#1,#2 \right)}
+\newcommand{\npairP}[2][n]{(#1,#2)}
 
@@
   cross-component function pointers) and can be efficient assuming
-  only \lars{can we say only here ?}
+  only
   one additional piece of processor support: an instruction for efficiently clearing
@@
     \hline
-    $\jmp{\lv}$&$\begin{multlined}
-                   \updateReg{\pcreg}{\\\updatePcPerm{\memreg(\lv)}}
-                 \end{multlined}$&\\
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$&$\var{newPc} = \updatePcPerm{\memreg(\lv)}$\\
     \hline
@@
 for which we receive a stack pointer in a register, $r_\stk$, that we reserve for this
-purpose. This stack
-pointer must be local and it must carry $\rwlx$ permission. It is crucial for
+purpose. This stack pointer must be local and it must carry $\rwlx$ permission. It is crucial for
 our approach that the stack is the only part of memory for which there will ever
@@
       \Regions & = \{\revoked\} \uplus \\
-                 &\hskip -2cm \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
-                 & \hskip -2cm \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
+                 &\;\quad \begin{multlined}
+                     \{\temp\} \times \States \times \Rels \times \\
+                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
+                   \end{multlined} \uplus \\
+                 &\;\quad \begin{multlined}
+                     \{\perma\} \times \States \times \Rels \times \\
+                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
+                   \end{multlined}
     \end{align*}
@@
   &\memSat{\ms}{W}
-    \text{ iff }
-    \left\{\begin{aligned}
-        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \\
-        &\quad (\hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\
+    \text{ iff }\\
+  &\quad \left\{\begin{aligned}
+        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\[-.2cm]
         &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
-        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \\
-        &\qquad \npair[n]{P(r)} \in H(s)(\xi^{-1}(W)))\\
+        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
       \end{aligned}\right.
@@
   \centering
+  \begin{withmathindent}{0cm}
     \begin{align*}
@@
       & \stdrr(W) \defeq \left\{ \npair{\reg} \; \middle| \;
-                  \begin{aligned}
-                      & \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
-                    & \;\;  \npair{\reg(r)} \in \stdvr(W) 
-                  \end{aligned} \right\} \\
-  & \\
+          \forall r \in \RegName \setminus \{\pcreg\} \ldotp \npair{\reg(r)}
+          \in \stdvr(W) \right\} \\
+      & \\[-.2cm]
       & \stder \; : \; \Worlds \nefun \UPred{\Words} \\
       & \stder(W) \defeq \left\{ \npair{\pc} \; \middle| \;
-                  \begin{aligned}
-                    & \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
-                    & \;\;\npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) 
-                  \end{aligned} \right\} \\
+        \begin{multlined}
+          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
+          \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W)
+        \end{multlined} \right\} \\
       & \\
@@
       &\stdvr(W)\defeq  \\
-    &\;\;\begin{aligned}[t]
-      & \{ \npair{i} \mid i \in \ints \} 
+      &\;\;\begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \}
         \union \\
@@
         \union \\
-      & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
+        & \left\{
+          \begin{multlined}
+            \npair{\stdcap[(\readonly,\gl)] } \mid \\
+            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+          \end{multlined} \right\}
         \union \\
-      & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \} 
+        & \left\{
+          \begin{aligned}
+            &\npair{\stdcap[(\readwrite,\gl)] } \mid \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\start,\addrend)} \in
+            \writeCond{}(\iota^\nwl,\gl)(W)
+          \end{aligned} \right\} \union \\
+        & \left\{
+          \begin{aligned}
+            &\npair{\stdcap[(\readwritel,\gl)] } \mid \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\start,\addrend)} \in
+            \writeCond{}(\iota^\pwl,\gl)(W)
+          \end{aligned} \right\}
         \union \\
-      & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
+        & \left\{
+          \begin{aligned}
+            &\npair{\stdcap[(\exec,\gl)]} \mid \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{aligned} \right\}
         \union \\
-      & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-      & \quad \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \} 
+        & \left\{
+          \begin{multlined}
+            \npair{\stdcap[(\entry,\gl)]} \mid \\
+            \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)
+          \end{multlined} \right\}
         \union \\
-      & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
+        & \left\{
+          \begin{aligned}
+            &\npair{\stdcap[(\rwx,\gl)]} \mid \\
+            &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+            &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
+            &\quad\npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{aligned} \right\}
         \union \\
-      & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
-      & \quad \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) 
-      \union \\
-      & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
-      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
-      & \quad \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \}
+        & \left\{
+          \begin{aligned}
+            &\npair{\stdcap[(\rwlx,\gl)]} \mid \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
+             &\quad\npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in
+            \execCond{}(\gl)(W)
+          \end{aligned} \right\}
       \end{aligned}
     \end{align*}
+  \end{withmathindent}
   \caption{Logical Relation}
@@
   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \; \middle| \;
-    \begin{aligned}
-      &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
-      &\forall \addr \in [\start,\addrend] \ldotp \\
-      &\;\; \nonlocal{\ms(\addr)} \text{\footnotesize{ and }}\\ 
-      &\;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-    \end{aligned}
-        \right\}
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \\
+    &\quad\left\{\npair{\hs} \; \middle| \;
+    \begin{multlined}
+      n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \forall \addr \in [\start,\addrend] \ldotp \\
+       {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }} \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
+    \end{multlined} \right\}
 \end{align*}
@@
       & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-      & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
+      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
     \end{aligned} \right\}\\
   & \writeCond{}(\iota,\gl)(W) =  \\
-  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
+  & \left\{
+    \npair{(\start,\addrend)}
+    \; \middle| \;
     \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
-      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and }}\\
-      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
+      & \exists r \in \var{localityReg}(g,W) \ldotp  W(r) \text{\footnotesize{ is address-stratified and}} \\
+      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
     \end{aligned} \right\}\\
   & \execCond{}(\gl)(W) = \\
-  & \left\{ \npair{(\var{perms},\start,\addrend)} \; \middle| \;
+  & \left\{
+    \begin{multlined}
+      \npairP{(\var{perms}, \\
+        \start,\addrend)}
+    \end{multlined}
+     \; \middle| \;
     \begin{aligned}
-      & \forall n' < n, W' \future W \ldotp \\
-      & \;\; \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
+      & \forall n' < n, W' \future W \ldotp \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
       & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
@@
       \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
-      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
-      \dom(\ms) = \dom(\ms') \wedge \\
+      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow 
+      \dom(\ms) = \dom(\ms') \text{\footnotesize{ and }} \\
       \forall \addr \in
@@
         \begin{aligned}[t]
-        &\perm = \exec \text{\footnotesize{ and }}\\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
       \end{aligned} \\
@@
   \end{align*}
-  then
-  \[
-    \npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)
-  \]
+  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
 \end{theorem}
@@
   \end{itemize}
-  if 
+  If 
   \[
@@
   \]
-  then
-  \[
-    m'(\flag) = 0
-  \]  
+  then $m'(\flag) = 0$
 \end{lemma}
@@
   if
-  $(\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m')$,
-  then 
   \[
-  m'(\flag) = 0.
+  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
   \]
+  then $m'(\flag) = 0$.
 \end{lemma}

### 2017-02-11 csf-2017/paper.tex
clean up of math in f1 lemma

@@
   Let
-  \begin{align*}
-    c_{\var{adv}} & \defeq ((\entry,\glob),\dots) \\
-    c_{f1} & \defeq ((\rwx,\glob),\dots) \\
-    c_\malloc & \defeq ((\entry,\glob),\dots) \\
-    c_{\var{stk}} & \defeq ((\rwlx,\local),\dots) \\
-    c_\link & \defeq ((\readonly,\glob),\dots)\\
-    \reg & \in \Regs \\
-    m & \defeq
-        \begin{aligned}[t]
-          &\hs_{f1} \uplus 
-        \hs_\flag \uplus                
-        \ms_{\var{link}} \uplus 
-        \hs_\adv \uplus \\
-        &\ms_{\malloc} \uplus 
-        \ms_{\var{stk}} \uplus
-        \ms_{\var{frame}}
-      \end{aligned}
-  \end{align*}
+\[
+    \begin{array}{r@{}lr@{}l}
+    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
+    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
+    c_\malloc &\; \defeq ((\entry,\glob),\dots) & & \\
+    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus} \\
+      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;} \ms_{\malloc} \uplus \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
+    \end{array}
+\]
   where each of the capability has an appropriate range of authority and pointer and
@@
   \end{itemize}
-  If 
-  \[
-    (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m'),
-  \]
+  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
   then $m'(\flag) = 0$

### 2017-02-11 csf-2017/paper.tex
more clean up

@@
     c_\malloc &\; \defeq ((\entry,\glob),\dots) & & \\
-    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus} \\
-      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;} \ms_{\malloc} \uplus \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
+    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
+      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
     \end{array}

### 2017-02-11 csf-2017/paper.tex
updates to loc/global cap figure

@@
   \includegraphics{w11}
-  \caption{Relation between local and global capabilities and temporary and permanent regions.}
+  \caption{Illustration of the relation between local and global capabilities and temporary and permanent regions.
+The table represents a memory and the colored fields are regions governing parts of the memory.
+Roughly speaking, global capabilities should never depend on memory governed by temporary regions.}
   \label{fig:cap-world}

### 2017-02-11 csf-2017/paper.tex
Fix a definition that didn't make formal sense (anymore).
Shorten a remark that was a bit overly technical.

@@
   \begin{array}{rcl}
-    \rv    &::=& n \mid \lv \\
+    \rv    &\in& \ints + \RegName \\
     i      &::=& 
@@
 
-The $\jmp{}$ instruction updates the program counter to a requested
-location, but it is complicated by the presence of \emph{enter
-  capabilities}, modeled after the
-M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
-cannot be used to read, write or execute and their current address and
-range cannot be modified. They can only be used to jump to, but when
-that happens, their permission is updated to $\exec$. This special
-behavior makes them suitable for representing a kind of closures: an
-opaque package containing a piece of code together with local
-encapsulated state. The idea is that one can build an enter capability
-$c = ((\entry,\gl),\start,\addrend,\addr)$ where the range
-$[\start,\addr-1]$ is used for storing the closure's local state (data
-or capabilities) and the range $[\addr,\addrend]$ can be used for
-storing instructions. The instructions and data are opaque to an
-adversary holding $c$ but when $c$ jumped to, the data becomes
-available to the closure's code through the updated version of $c$
-that is now in the $\pcreg$\footnote{Note that the capability in the
-  $\pcreg$-register can be moved to another register and used to load
-  from.}. \lau{As we talked about what is described here, is not how we make closures in the TR, so we should consider whether we should refrain from using that word here. }
+The $\jmp{}$ instruction updates the program counter to a requested location,
+but it is complicated by the presence of \emph{enter capabilities}, modeled
+after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
+cannot be used to read, write or execute and their current address and range
+cannot be modified. They can only be used to jump to, but when that happens,
+their permission is updated to $\exec$. This special behavior makes them
+suitable for representing a kind of closures: an opaque package containing a
+piece of code together with local encapsulated state. The idea is that one can
+build an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where the
+range $[\start,\addr-1]$ is used for storing the closure's local state (data or
+capabilities) and the range $[\addr,\addrend]$ can be used for storing
+instructions. The instructions and data are opaque to an adversary holding $c$
+but when $c$ jumped to, the data becomes available to the closure's code through
+the updated version of $c$ that is now in the $\pcreg$\footnote{Note that the
+  capability in the $\pcreg$-register can be used to load from.}. \lau{As we
+  talked about what is described here, is not how we make closures in the TR, so
+  we should consider whether we should refrain from using that word here. }
 

### 2017-02-11 csf-2017/paper.tex
Start thinking about Related Work.

@@
 
+Cap-mach:
 \begin{itemize}
-\item Dreyer-Neis-Birkedal
-\item CHERI papers
-\item Akram's thesis
-\item \url{http://2016.splashcon.org/event/splash-2016-oopsla-gentrification-gone-too-far-affordable-2nd-class-values-for-fun-and-co-effect}
-\item other papers that enforce well-bracketed control flow at low level
-(using a trusted stack manager)
+\item \cite{Watson2015Cheri} CHERI papers
+\item \cite{akram_el-korashy_formal_2016}
+\end{itemize}
+
+Logical relation:
 \begin{itemize}
-\item \url{http://ieeexplore.ieee.org/abstract/document/7536364/}
-\item \url{http://ieeexplore.ieee.org/abstract/document/7536366/}
+\item \cite{Hur:2011:KLR:1926385.1926402}
+\item \cite{Dreyer:jfp12}
+\item \cite{Devriese:2016ObjCap}
+\item \cite{osvald_gentrification_2016}
+\end{itemize}
+
+Other papers that enforce well-bracketed control flow at low level (using a
+trusted stack manager)
+\begin{itemize}
+\item \cite{patrignani_modular_2016-1}
+\item \cite{juglaret_beyond_2016-1}
 \item other stuff?
-\end{itemize}
 \end{itemize}

### 2017-02-13 csf-2017/paper.tex
Removed dot in title
Pass over examples section

@@
 \title{Reasoning about a Capability Machine with Local Capabilities\\
- Provably Safe Stack and Return Pointer Management (without OS Support).}
+ Provably Safe Stack and Return Pointer Management (without OS Support)}
 
@@
 \section{Stack and return pointer management using local capabilities}
+\label{sec:stack-and-return-pointer}
 \lau{Do we want to use the term ``pointer'' at any point? I have also used it, but I guess it might create some confusing to whether we have traditional pointer as well as capabilities?}
@@
 received upon the second invocation. \lau{If we need space, then maybe one of the examples is sufficient.}
+\lau{In the examples, we refer to the local return pointer as
+  ``protected return pointer''. We should either introduce that here
+  or use a different formulation in the examples. }
 
@@
 safety allows us to prove local state encapsulation and control-flow
-integrity properties for challenging examples. 
-Since we are dealing with assembly language, 
-there are many details to the formal treatment , and therefore we 
-necessarily have to omit some details
-in the lemmas we state here.  The interested reader can find all the
+integrity properties for challenging program examples.  Since we are
+dealing with assembly language, there are many details to the formal
+treatment, and therefore we necessarily have to omit some details in
+the lemmas we state here.  The interested reader can find all the
 technical details in the techincal report~\cite{XXX}.
@@
 The capability machine provides encapsulation of local state. The
-programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and 
-\texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.
-The two programs \texttt{f1} and \texttt{f2} are very similar. They both
-store some local data, call an untrusted piece of code ($\adv$), and
-then tests whether local state is unchanged. They differ in the way
-they do this. Program \texttt{f1} assumes it has access to a stack capability
+programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
+\figurename~\ref{fig:prog-f2} demonstrate this.  The two programs
+\texttt{f1} and \texttt{f2} are very similar. They both store some
+local state, call an untrusted piece of code ($\adv$), and then tests
+whether the local state is unchanged. They differ in the way they do
+this. Program \texttt{f1} assumes it has access to a stack capability
 and uses the stack to store its local state. It also uses the
-stack-based calling convention to call the untrusted code. On the
+stack-based calling convention \texttt{scall} to call the untrusted code. On the
 other hand, \texttt{f2} allocates a new piece of memory for its local
@@
 For both programs, we can prove that if they are linked with an
-adversary, $\adv$, that is allowed to allocate new memory but otherwise has no
-capabilities, then the assertion will never fail when
+adversary, $\adv$, that is allowed to allocate new memory but
+otherwise has no capabilities, then the assertion will never fail when
 executing the program (see Lemmas~\ref{lem:correctness-f1}
-and~\ref{lem:correctness-f2} below).
-\lars{shouldn't we say already here that the adversary is a global 
-  capability and that is why, intuitively, the local state is
-  preserved ?}
-The two examples also illustrate
-the versatility of the logical relation. The two examples use
-different calling convention, but the logical relation is not specific
-to any calling convention, so we can use it to reason about both
-programs.
+and~\ref{lem:correctness-f2} below).  \lars{shouldn't we say already
+here that the adversary is a global capability and that is why,
+intuitively, the local state is preserved ?} \lau{Due to the ``order
+of control'', \texttt{f1} runs first. The lemma is quite restrictive
+on the adversary in the sense that it only has access to malloc (and
+whatever we give et access to). Specifically, it does not have access
+to the stack capability and because of that we don't need to check it
+first. We could write that the lemmas are limited in their assumptions
+to make them simpler and a more complicated example is coming up.}
+The two examples also illustrate the versatility of the logical
+relation. The two examples use different calling convention, but the
+logical relation is not specific to any calling convention, so we can
+use it to reason about both programs.
 
@@
 need a way to observe whether the assertion fails. To this end, we
-assume access to a flag (an address in memory)
-which will be set to $1$ by the assertion if it
-fails. A slightly simplified version of the correctness lemma for
-\texttt{f1} is:
+assume access to a flag (an address in memory). When the assertion
+fails, it sets the flag to $1$ and halts. A slightly simplified
+version of the correctness lemma for \texttt{f1} is:
 \begin{lemma}
@@
 
-To prove Lemma~\ref{lem:correctness-f1}, it suffices to show
-that the start configuration is capability safe (in the $\observations$ relation) for a
-world with a permanent region that requires the assertion flag to be
-0. This is done using an anti-reduction lemma which states that for a
-configuration to be capability safe, it is sufficient to
-show that it is in it at a later point of the execution. To reason
-about \texttt{scall}, it suffices to show that (1) the configuration just
-after we jumped is capability safe and (2) that the configuration just
-after \texttt{scall} is done cleaning up is capability safe.
-We use the Fundamental Theorem to reason about the unknown code, but
-notice that the Fundamental Theorem says nothing about enter
-capabilities and that the adversary capability is an enter
+To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the
+start configuration is capability safe (in the $\observations$
+relation) for a world with a permanent region that requires the
+assertion flag to be 0. This is done using an anti-reduction lemma
+which states that for a configuration to be capability safe, it is
+sufficient to show that it is in it at a later point of the
+execution. To reason about \texttt{scall}, it suffices to show that
+(1) the configuration just after we jumped is capability safe and (2)
+that the configuration just after \texttt{scall} is done cleaning up
+is capability safe.  We use the Fundamental Theorem to reason about
+the unknown code, but notice that the Fundamental Theorem says nothing
+about enter capabilities and that the adversary capability is an enter
 capability. Luckily the enter capability becomes a read/execute
@@
 
+\lau{in the following paragraph, we could make a reference to
+  \textbf{invoking the adversary} from
+  section~\ref{sec:stack-and-return-pointer} where the problem is
+  first introduced. In fact, this paragraph is a reiteration of the
+  problem in the aformentioned section.}
 In \texttt{f3} there are two calls to an adversary and in order for
@@
 well-bracketed. If the adversary is able to store the protected return
-pointer, so it is available for the second call, then it can be used
-in a context where \texttt{f3} has 2 on top of its stack which will cause the
+pointer from the first call, so it is available for the second call, then it can be used
+in a context where \texttt{f3} has $2$ on top of its stack which will cause the
 assertion to fail. The protected return pointer is passed as a local
@@
 the adversary jumps to the protected return pointer, \texttt{f3} can
-revoke it (by overwritting it which happens as part of the
+revoke it (by erasing it which happens as part of the
 \texttt{scall}). In other words, the adversary cannot store the
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
+  Lemma~\ref{lem:correctness-f1}
   if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
+  Lemma~\ref{lem:correctness-f1} 
   if

### 2017-02-13 csf-2017/paper.tex
Examples section

@@
 safety allows us to prove local state encapsulation and control-flow
-integrity properties for challenging program examples.  Since we are
-dealing with assembly language, there are many details to the formal
-treatment, and therefore we necessarily have to omit some details in
-the lemmas we state here.  The interested reader can find all the
-technical details in the techincal report~\cite{XXX}.
+integrity properties for challenging program examples. The security
+measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
+ensure these properties. Since we are dealing with assembly language,
+there are many details to the formal treatment, and therefore we
+necessarily have to omit some details in the lemmas we state here.
+The interested reader can find all the technical details in the
+techincal report~\cite{XXX}.
 
@@
 litterature\cite{XXX}. It consists of two parts, \texttt{g1} and
-\texttt{f4}. \texttt{g1} is a closure generator that generates a
-closure with one variable $x$ set to 0 in its environment and \texttt{f4} as
-the program. The program \texttt{f4} expects to arguments: a stack and
-a callback. The stack is necessary to use \texttt{scall}, but because
-it comes from an untrusted source, we need to make sure that it has
-the correct permission and fits with our stack dicipline. The callback
-must be global for the program to work, we will get back to this
-point. The program now sets $x$ to 0 and calls the callback. When it
-returns from the callback, it sets $x$ to 1 and calls the callback a
-second time. When it returns from the callback, it asserts $x$ is 1
-and returns.
+\texttt{f4}. \texttt{g1} is a closure generator that generates
+closures with one variable $x$ set to $0$ in its environment and
+\texttt{f4} as the program. The program \texttt{f4} expects two
+arguments: a stack and a callback. The stack is necessary to use
+\texttt{scall}, but because it comes from an untrusted source, we need
+to make sure that it has the correct permission and fits with our
+stack dicipline. The callback must be global for the program to work,
+we will get back to this point. The program now sets $x$ to $0$ and
+calls the callback. When it returns from the callback, it sets $x$ to
+$1$ and calls the callback a second time. When it returns from the
+callback, it asserts $x$ is $1$ and returns.
 
@@
 the second callback and use it in the first, then the assertion will
-fail because $x$ is 0 before the first callback. The protected return
+fail because $x$ is $0$ before the first callback. The protected return
 pointer cannot be transfered between the callbacks, so this does not

### 2017-02-13 csf-2017/paper.tex
Lemma 1 added assumption

@@
     c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
-    c_\malloc &\; \defeq ((\entry,\glob),\dots) & & \\
+    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
     m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\

### 2017-02-13 csf-2017/paper.tex
updated g1 text and added a figure

@@
 \usepackage{tikz}
+\usetikzlibrary{positioning}
 \usepackage{multirow,bigdelim}
@@
 \end{figure}
-
+\lau{Add something about the correctness lemma? Not sure if I have
+  anything important to say.}
 \begin{lemma}
@@
 callback, it asserts $x$ is $1$ and returns.
-
-To illustrate how subtle this program is, consider one way an
-adversary may try to make the assertion fail. In the second callback
-an adversary can get to the first callback by calling the closure one
-more time. If the adversary can save the protected return pointer in
-the second callback and use it in the first, then the assertion will
-fail because $x$ is $0$ before the first callback. The protected return
-pointer cannot be transfered between the callbacks, so this does not
-work.
-
-\begin{lemma}
-  \label{lem:correctness-g1}
-  Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} 
-  if
-  \[
-  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
-  \]
-  then $m'(\flag) = 0$.
-\end{lemma}
-
-As mentioned, the callback must be global. This is to make sure the
-callback is not a piece of code on the stack that restores the
-adversaries original stack capability breaking the encapsulation of
-our local stack frame. In the proof of Lemma~\ref{lem:correctness-g1},
-this also shows up. When we reason about \texttt{f4}, we revoke all
-the local state of the adversary which means that we consider a
-private future world. We use the $\execCond{}$ to reason about the
-jump to the callback and because the callback is global, it works for
-private future worlds. Had the callback been local, then the
-$\execCond{}$ would only work for public future worlds, and we would
-be stuck in the proof.
-
-In Lemma~\ref{lem:correctness-g1} the order of control has been
-inverted compared to the previous lemmas. In this lemma, the adversary
-assumes control first with a capability for the closure creator
-\texttt{g1}. This means we need to check all arguments are safe
-to use and that we clean up before returning in the end. The inversion
-of control poses an interesting challenge when it comes to reasoning
-bout the adversaries local state during the execution of
-\texttt{f4}. During the execution of \texttt{f4} and the callbacks the
-adversary should not be able to rely on its local state from outside
-the call of \texttt{f4}. This is easily done by revoking all the
-temporary regions of the world given at the start of \texttt{f4}. The
-real challenge comes when \texttt{f4} returns and needs to guarantee
-that the protocols of the adversaries local state has not been
-violated. To this end, we need to reinstate regions we revoked which
-is done using the masking regions.
 
@@
 
+To illustrate how subtle this program is, consider one way an
+adversary may try to make the assertion fail. In the second callback
+an adversary can get to the first callback by calling the closure one
+more time. If the adversary can save the protected return pointer in
+the second callback and use it in the first, then the assertion will
+fail because $x$ is $0$ before the first callback. The protected return
+pointer cannot be transfered between the callbacks, so this does not
+work.
+
+\begin{lemma}
+  \label{lem:correctness-g1}
+  Making similar assumptions about capabilities and linking as in
+  Lemma~\ref{lem:correctness-f1} 
+  if
+  \[
+  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
+  \]
+  then $m'(\flag) = 0$.
+\end{lemma}
+
+As mentioned, the callback must be global. This is to make sure the
+callback is not a piece of code on the stack that restores the
+adversaries original stack capability breaking the encapsulation of
+our local stack frame. In the proof of Lemma~\ref{lem:correctness-g1},
+this also shows up. When we reason about \texttt{f4}, we revoke all
+the local state of the adversary which means that we consider a
+private future world. We use the $\execCond{}$ to reason about the
+jump to the callback and because the callback is global, it works for
+private future worlds. Had the callback been local, then the
+$\execCond{}$ would only work for public future worlds, and we would
+be stuck in the proof.
+
+In Lemma~\ref{lem:correctness-g1} the order of control has been
+inverted compared to the previous lemmas. In this lemma, the adversary
+assumes control first with a capability for the closure creator
+\texttt{g1}. This means we need to check all arguments are safe
+to use and that we clean up before returning in the end. The inversion
+of control poses an interesting challenge when it comes to reasoning
+about the adversaries local state during the execution of
+\texttt{f4}. During the execution of \texttt{f4} and the callbacks, the
+adversary should not be able to rely on its local state from before
+the call of \texttt{f4}. This is easily done by revoking all the
+temporary regions of the world given at the start of \texttt{f4}. The
+real challenge comes when \texttt{f4} returns and we need to establish
+that the protocols of the adversaries local state has not been
+violated. This can be done by reinstating all of the adverseries old
+regions, but we can only do this if the old adverserial regions are
+not masked by permanent regions. As illustrated in
+\figurename~\ref{fig:worlds-in-pf-g1}, we know this must be the
+case. Whenever we are given a world, we get a public future world
+where it is not allowed to use permanent regions for masking and in
+the private future worlds we construct, we know it is not done
+either. The return pointer \texttt{f4} gets from the adversary is
+valid in future worlds of $W_1$, so in order to use it to argue it is
+safe to return $W_6$ needs to be a public future worlds of $W_1$. We
+can construct this as a private future world of $W_5$, but only
+because we can reinstate the temporary regions of $W_1$ with the same
+region names.
+
+\begin{figure}[hptb]
+  \centering
+\begin{tikzpicture}[main node/.style={}, node distance=1.5cm]
+  \node[main node] (1) {$W_1$};
+  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f3} called}}};
+  \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};
+
+  \node[main node] (2) [above right of=1]{$W_2$};
+  \node[main node] (label2) [above of=2,yshift=-1cm] {\textit{\footnotesize{first callback}}};
+
+  \node[main node,left] (const) [above of=given,yshift=-0.4cm] {\parbox{1.5cm}{\footnotesize{Constructed:}}};
+
+  \node[main node] (3) [below right of=2]{$W_3$};
+  \node[main node] (label3) [below of=3,yshift=0.75cm] {\footnotesize{\textit{callback returns}}};
+
+  \node[main node] (4) [above right of=3]{$W_4$};
+  \node[main node] (label4) [above of=4,yshift=-0.75cm] {\footnotesize{\textit{second callback}}};
+
+  \node[main node] (5) [below right of=4]{$W_5$};
+  \node[main node] (label5) [below of=5,yshift=1cm] {\footnotesize{\textit{callback returns}}};
+
+  \node[main node] (6) [above right of=5]{$W_6$};
+  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
+  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
+       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
+       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
+       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
+       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
+\end{tikzpicture}
+  \caption{Worlds in the proof of Lemma~\ref{lem:correctness-g1}}
+  \label{fig:worlds-in-pf-g1}
+\end{figure}
 \begin{itemize}

### 2017-02-13 csf-2017/paper.tex
updates to the new figure
Rewrote some more about g1

@@
 the call of \texttt{f4}. This is easily done by revoking all the
-temporary regions of the world given at the start of \texttt{f4}. The
-real challenge comes when \texttt{f4} returns and we need to establish
-that the protocols of the adversaries local state has not been
-violated. This can be done by reinstating all of the adverseries old
-regions, but we can only do this if the old adverserial regions are
-not masked by permanent regions. As illustrated in
-\figurename~\ref{fig:worlds-in-pf-g1}, we know this must be the
-case. Whenever we are given a world, we get a public future world
-where it is not allowed to use permanent regions for masking and in
-the private future worlds we construct, we know it is not done
-either. The return pointer \texttt{f4} gets from the adversary is
-valid in future worlds of $W_1$, so in order to use it to argue it is
-safe to return $W_6$ needs to be a public future worlds of $W_1$. We
-can construct this as a private future world of $W_5$, but only
-because we can reinstate the temporary regions of $W_1$ with the same
-region names.
-
-\begin{figure}[hptb]
+temporary regions of the world given at the start of \texttt{f4}. When
+\texttt{f4} returns, then the adversary is allowed to rely on its
+local state again which means that we need to guarantee that the local
+state is unchanged. To illustrate how this is done in the proof of
+Lemma~\ref{lem:correctness-g1}, consider
+\figurename~\ref{fig:worlds-in-pf-g1}. The return pointer given to
+\texttt{f4} might be local and it is capability safe in $W_1$, so it
+can only be used in public future worlds. Let us see how $W_6$ is
+constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed by revoking all
+temporary regions and adding a $\iota^\pwl$ region for the stack we
+pass away and static temporary regions (a region that accepts exactly
+one memory) for our local stack as well as the adversaries local stack.
+The given worlds $W_3$ and $W_5$ are public future worlds. None of
+these worlds can have masked the temporary regions of $W_1$ with permanent
+regions. $W_6$ is constructed by revoking the temporary regions of
+$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$
+a public future world of $W_1$. It is therefore safe to use the return
+pointer to return to the adversary.
+
+\begin{figure}[t]
   \centering
@@
 \end{tikzpicture}
-  \caption{Worlds in the proof of Lemma~\ref{lem:correctness-g1}}
+  \caption{Illustration of some of the worlds in the proof of
+    Lemma~\ref{lem:correctness-g1}. The first row of worlds are worlds
+  we construct during the proof. The second row of worlds are worlds
+  we are given during the proof. $W_6$ is constructed such that $W_6
+  \futurewk W_1$.}
   \label{fig:worlds-in-pf-g1}

### 2017-02-13 csf-2017/paper.tex
Added question to example section
Math update
Work on related work.

--- 2017-02-13 csf-2017/paper.tex
@@
 \title{Reasoning about a Capability Machine with Local Capabilities\\
- Provably Safe Stack and Return Pointer Management (without OS Support)}
+ Provably Safe Stack and Return Pointer Management (without OS Support).}
 
@@
 \section{A capability machine with local capabilities}
+\label{sec:capab-mach-with}
 In this paper, we work with a formal capability machine that has all the
@@
 
+We point out that our local capabilities capture only a part of the semantics of
+local capabilities in CHERI. Specifically, on top of the above restrictions,
+CHERI's default implementation of CCall additionally forbids local capabilities
+from being passed accross module boundaries, something which we rely on
+crucially in this paper, for passing around local return pointers and stack
+capabilities. However, CHERI's CCall is not implemented in hardware, but in
+software, precisely to allow experimenting with alternative models like ours.
+
 % Contrary to more traditional assembly languages, capability machines do not
@@
 \section{Stack and return pointer management using local capabilities}
-\label{sec:stack-and-return-pointer}
 \lau{Do we want to use the term ``pointer'' at any point? I have also used it, but I guess it might create some confusing to whether we have traditional pointer as well as capabilities?}
@@
 received upon the second invocation. \lau{If we need space, then maybe one of the examples is sufficient.}
-\lau{In the examples, we refer to the local return pointer as
-  ``protected return pointer''. We should either introduce that here
-  or use a different formulation in the examples. }
 
@@
 \section{Logical relation}
+\label{sec:logical-relation}
 In order to be able to reason about programs running on the capability
@@
 safety allows us to prove local state encapsulation and control-flow
-integrity properties for challenging program examples. The security
-measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
-ensure these properties. Since we are dealing with assembly language,
-there are many details to the formal treatment, and therefore we
-necessarily have to omit some details in the lemmas we state here.
-The interested reader can find all the technical details in the
-techincal report~\cite{XXX}.
+integrity properties for challenging examples. 
+Since we are dealing with assembly language, 
+there are many details to the formal treatment , and therefore we 
+necessarily have to omit some details
+in the lemmas we state here.  The interested reader can find all the
+technical details in the techincal report~\cite{XXX}.
 
@@
 The capability machine provides encapsulation of local state. The
-programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
-\figurename~\ref{fig:prog-f2} demonstrate this.  The two programs
-\texttt{f1} and \texttt{f2} are very similar. They both store some
-local state, call an untrusted piece of code ($\adv$), and then tests
-whether the local state is unchanged. They differ in the way they do
-this. Program \texttt{f1} assumes it has access to a stack capability
+programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and 
+\texttt{f2} \figurename~\ref{fig:prog-f2} demonstrate this.
+The two programs \texttt{f1} and \texttt{f2} are very similar. They both
+store some local data, call an untrusted piece of code ($\adv$), and
+then tests whether local state is unchanged. They differ in the way
+they do this. Program \texttt{f1} assumes it has access to a stack capability
 and uses the stack to store its local state. It also uses the
-stack-based calling convention \texttt{scall} to call the untrusted code. On the
+stack-based calling convention to call the untrusted code. On the
 other hand, \texttt{f2} allocates a new piece of memory for its local
@@
 For both programs, we can prove that if they are linked with an
-adversary, $\adv$, that is allowed to allocate new memory but
-otherwise has no capabilities, then the assertion will never fail when
+adversary, $\adv$, that is allowed to allocate new memory but otherwise has no
+capabilities, then the assertion will never fail when
 executing the program (see Lemmas~\ref{lem:correctness-f1}
-and~\ref{lem:correctness-f2} below).  \lars{shouldn't we say already
-here that the adversary is a global capability and that is why,
-intuitively, the local state is preserved ?} \lau{Due to the ``order
-of control'', \texttt{f1} runs first. The lemma is quite restrictive
-on the adversary in the sense that it only has access to malloc (and
-whatever we give et access to). Specifically, it does not have access
-to the stack capability and because of that we don't need to check it
-first. We could write that the lemmas are limited in their assumptions
-to make them simpler and a more complicated example is coming up.}
-The two examples also illustrate the versatility of the logical
-relation. The two examples use different calling convention, but the
-logical relation is not specific to any calling convention, so we can
-use it to reason about both programs.
+and~\ref{lem:correctness-f2} below).
+\lars{shouldn't we say already here that the adversary is a global 
+  capability and that is why, intuitively, the local state is
+  preserved ?}
+The two examples also illustrate
+the versatility of the logical relation. The two examples use
+different calling convention, but the logical relation is not specific
+to any calling convention, so we can use it to reason about both
+programs.
 
@@
 need a way to observe whether the assertion fails. To this end, we
-assume access to a flag (an address in memory). When the assertion
-fails, it sets the flag to $1$ and halts. A slightly simplified
-version of the correctness lemma for \texttt{f1} is:
+assume access to a flag (an address in memory)
+which will be set to $1$ by the assertion if it
+fails. A slightly simplified version of the correctness lemma for
+\texttt{f1} is:
 \begin{lemma}
@@
     c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
-    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
+    c_\malloc &\; \defeq ((\entry,\glob),\dots) & & \\
     m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
@@
 
-To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the
-start configuration is capability safe (in the $\observations$
-relation) for a world with a permanent region that requires the
-assertion flag to be 0. This is done using an anti-reduction lemma
-which states that for a configuration to be capability safe, it is
-sufficient to show that it is in it at a later point of the
-execution. To reason about \texttt{scall}, it suffices to show that
-(1) the configuration just after we jumped is capability safe and (2)
-that the configuration just after \texttt{scall} is done cleaning up
-is capability safe.  We use the Fundamental Theorem to reason about
-the unknown code, but notice that the Fundamental Theorem says nothing
-about enter capabilities and that the adversary capability is an enter
+To prove Lemma~\ref{lem:correctness-f1}, it suffices to show
+that the start configuration is capability safe (in the $\observations$ relation) for a
+world with a permanent region that requires the assertion flag to be
+0. This is done using an anti-reduction lemma which states that for a
+configuration to be capability safe, it is sufficient to
+show that it is in it at a later point of the execution. To reason
+about \texttt{scall}, it suffices to show that (1) the configuration just
+after we jumped is capability safe and (2) that the configuration just
+after \texttt{scall} is done cleaning up is capability safe.
+We use the Fundamental Theorem to reason about the unknown code, but
+notice that the Fundamental Theorem says nothing about enter
+capabilities and that the adversary capability is an enter
 capability. Luckily the enter capability becomes a read/execute
@@
 
-\lau{in the following paragraph, we could make a reference to
-  \textbf{invoking the adversary} from
-  section~\ref{sec:stack-and-return-pointer} where the problem is
-  first introduced. In fact, this paragraph is a reiteration of the
-  problem in the aformentioned section.}
 In \texttt{f3} there are two calls to an adversary and in order for
@@
 well-bracketed. If the adversary is able to store the protected return
-pointer from the first call, so it is available for the second call, then it can be used
-in a context where \texttt{f3} has $2$ on top of its stack which will cause the
+pointer, so it is available for the second call, then it can be used
+in a context where \texttt{f3} has 2 on top of its stack which will cause the
 assertion to fail. The protected return pointer is passed as a local
@@
 the adversary jumps to the protected return pointer, \texttt{f3} can
-revoke it (by erasing it which happens as part of the
+revoke it (by overwritting it which happens as part of the
 \texttt{scall}). In other words, the adversary cannot store the
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1}
+  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
   if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
@@
 litterature\cite{XXX}. It consists of two parts, \texttt{g1} and
-\texttt{f4}. \texttt{g1} is a closure generator that generates
-closures with one variable $x$ set to $0$ in its environment and
-\texttt{f4} as the program. The program \texttt{f4} expects two
-arguments: a stack and a callback. The stack is necessary to use
-\texttt{scall}, but because it comes from an untrusted source, we need
-to make sure that it has the correct permission and fits with our
-stack dicipline. The callback must be global for the program to work,
-we will get back to this point. The program now sets $x$ to $0$ and
-calls the callback. When it returns from the callback, it sets $x$ to
-$1$ and calls the callback a second time. When it returns from the
-callback, it asserts $x$ is $1$ and returns.
+\texttt{f4}. \texttt{g1} is a closure generator that generates a
+closure with one variable $x$ set to 0 in its environment and \texttt{f4} as
+the program. The program \texttt{f4} expects to arguments: a stack and
+a callback. The stack is necessary to use \texttt{scall}, but because
+it comes from an untrusted source, we need to make sure that it has
+the correct permission and fits with our stack dicipline. The callback
+must be global for the program to work, we will get back to this
+point. The program now sets $x$ to 0 and calls the callback. When it
+returns from the callback, it sets $x$ to 1 and calls the callback a
+second time. When it returns from the callback, it asserts $x$ is 1
+and returns.
 
@@
 the second callback and use it in the first, then the assertion will
-fail because $x$ is $0$ before the first callback. The protected return
+fail because $x$ is 0 before the first callback. The protected return
 pointer cannot be transfered between the callbacks, so this does not
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} 
+  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
   if
@@
 
-Cap-mach:
-\begin{itemize}
-\item \cite{Watson2015Cheri} CHERI papers
-\item \cite{akram_el-korashy_formal_2016}
-\end{itemize}
-
-Logical relation:
-\begin{itemize}
-\item \cite{Hur:2011:KLR:1926385.1926402}
-\item \cite{Dreyer:jfp12}
-\item \cite{Devriese:2016ObjCap}
-\item \cite{osvald_gentrification_2016}
-\end{itemize}
+Capability machines originate with Dennis and Van
+Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
+\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview of
+previous work in the area. The capability machine formalised in
+Section~\ref{sec:capab-mach-with} is a simple but representative model, modeled
+mainly after the M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter
+pointers resemble the M-Machine's) and
+CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the memory
+capabilities resemble CHERI's). The latter is a recent and relatively mature
+capability machine, which combines capabilities with a virtual memory approach,
+in the interest of backwards compatibility and gradual adoption. Our local
+capabilities are inspired by CHERI's, although as discussed, they can be passed
+accross module boundaries, contrary to what is enforced by CHERI's default CCall
+implementation.
+
+We believe local capabilities
 
@@
 \end{itemize}
+CFI: 
+\begin{itemize}
+\item \cite{budiu_architectural_2006,abadi_control-flow_2005}
+\end{itemize}
+
+Our logical relation is a unary, step-indexed Kripke logical relation (SKLR)
+with recursive
+worlds~\cite{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
+closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
+capability-safety in a high-level JavaScript-like lambda calculus. Our
+fundamental theorem is similar to theirs and can be seen as a formulation of
+capability-safety of the capability-machine. Because we are not interested in an
+externally observable notion of side-effects (like console output or memory
+access traces), we do not require their notion of effect parametricity.
+
+The idea of applying SKLRs to an assembly language was applied before by
+\citet{Hur:2011:KLR:1926385.1926402}, who constructed a binary SKLR for
+formulating and proving correctness of an ML-to-assembly compiler. Their
+condition on the assembly code are quite loose (for example, they allow
+hand-written and self-modifying assmebly code), but they do assume code that it
+is non-adversarial and respects the ABI of the high-level language (for example,
+the code does not store return pointers and invoke them after their intended
+lifetime). As a result, they can treat return pointers as evaluation contexts in
+a lambda calculus and use biorthogonality~\cite{pitts_higher_1998}. This means
+that they define the logical relation to accept evaluation contexts if they
+produce related observations for related values and to accept terms if they
+produce related observations when plugged into related evaluation contexts. In
+our setting, there is no guarantee that adversarial code will treat return
+pointers as intended, so, as described in Section~\ref{sec:logical-relation}, we
+define our logical in a (arguably natural but still) novel way, using a single
+orthogonal closure, i.e. accept terms if they produce related observations when
+given related arguments, where those arguments include the return pointer.
+
+Another novel aspect of our logical relation is how we model the temporary
+nature of local capabilities. The main insight is that the temporary, revokable
+nature of these capabilities can be seen as a generalisation of the special
+status of evaluation contexts in lambda calculi. Closely related here is a study
+by \citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and control
+effects on local relational reasoning. DNB formally capture this special status
+of evaluation contexts using Kripke worlds with public and private future world
+relations. Our public and private worlds are inspired by theirs and they are at
+times remarkably similar. For example, the state-transitition system in the
+proof of our assembly version of the awkward example, is exactly the same as the
+one used by DNB for the original awkward example. Nevertheless, there are also
+remarkable differences: in DNB's logical relation, public future worlds only
+turn up in the continuation relation (which we do not have, see above). This is
+different from our value relation which uses public and private future worlds
+depending on the locality of the capability at hand. Additionally, our worlds
+are a bit more complex because, to allow store-local capabilities, they can have
+temporary regions that are only monotonous w.r.t. public future worlds, while
+DNB's worlds are entirely permanent.
+
+As a side note: our model of local/global capabilities may also be applicable
+for reasoning about related ideas in higher-level languages. Specifically,
+\citet{osvald_gentrification_2016}'s second-class or local values in Scala are
+remarkably similar to the local capabilities we study and we expect they will
+support similar reasoning patterns.
+
+\begin{itemize}
+\item \cite{akram_el-korashy_formal_2016}
+\end{itemize}
 

### 2017-02-13 csf-2017/paper.tex

--- 2017-02-13 csf-2017/paper.tex
@@
     \begin{aligned}
-      & \forall n' < n, W' \future W \ldotp \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
+      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
       & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
@@
 
+\lau{Add that the difficulties of the high-level version is also
+  present here nad solved in the same way?}
 \begin{figure}[t]

### 2017-02-13 csf-2017/paper.tex
Two bibliography fixes.

@@
 lifetime). As a result, they can treat return pointers as evaluation contexts in
-a lambda calculus and use biorthogonality~\cite{pitts_higher_1998}. This means
-that they define the logical relation to accept evaluation contexts if they
-produce related observations for related values and to accept terms if they
+a lambda calculus and use biorthogonality~\cite{pitts_operational_1998}. This
+means that they define the logical relation to accept evaluation contexts if
+they produce related observations for related values and to accept terms if they
 produce related observations when plugged into related evaluation contexts. In
@@
 nature of these capabilities can be seen as a generalisation of the special
-status of evaluation contexts in lambda calculi. Closely related here is a study
-by \citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and control
-effects on local relational reasoning. DNB formally capture this special status
-of evaluation contexts using Kripke worlds with public and private future world
-relations. Our public and private worlds are inspired by theirs and they are at
-times remarkably similar. For example, the state-transitition system in the
-proof of our assembly version of the awkward example, is exactly the same as the
-one used by DNB for the original awkward example. Nevertheless, there are also
-remarkable differences: in DNB's logical relation, public future worlds only
-turn up in the continuation relation (which we do not have, see above). This is
-different from our value relation which uses public and private future worlds
-depending on the locality of the capability at hand. Additionally, our worlds
-are a bit more complex because, to allow store-local capabilities, they can have
-temporary regions that are only monotonous w.r.t. public future worlds, while
-DNB's worlds are entirely permanent.
+status of evaluation contexts in lambda calculi. Closely related to this is a
+study by \citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and
+control effects on local relational reasoning. DNB formally capture this special
+status of evaluation contexts using Kripke worlds with public and private future
+world relations. Our public and private worlds are inspired by theirs and they
+are at times remarkably similar. For example, the state-transitition system in
+the proof of our assembly version of the awkward example, is exactly the same as
+the one used by DNB for the original awkward example. Nevertheless, there are
+also remarkable differences: in DNB's logical relation, public future worlds
+only turn up in the continuation relation (which we do not have, see above).
+This is different from our value relation which uses public and private future
+worlds depending on the locality of the capability at hand. Additionally, our
+worlds are a bit more complex because, to allow store-local capabilities, they
+can have temporary regions that are only monotonous w.r.t. public future worlds,
+while DNB's worlds are entirely permanent.
 
@@
 %\bibliographystyle{IEEEtran}
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{IEEEabrv,references}

### 2017-02-13 csf-2017/paper.tex
bibstyle from plain to plainnat
Related work typos

@@
 condition on the assembly code are quite loose (for example, they allow
-hand-written and self-modifying assmebly code), but they do assume code that it
+hand-written and self-modifying assmebly code), but they do assume code that
 is non-adversarial and respects the ABI of the high-level language (for example,
@@
 pointers as intended, so, as described in Section~\ref{sec:logical-relation}, we
-define our logical in a (arguably natural but still) novel way, using a single
+define our logical relation in a (arguably natural but still) novel way, using a single
 orthogonal closure, i.e. accept terms if they produce related observations when

### 2017-02-13 csf-2017/paper.tex
put list of possible examples in a comment
Try to make a sentence compute...

@@
 The basic idea of our approach is quite simple: we stick to a single, rather
-standard C stack and register-passed return pointers in much the same way as a
-standard C ABI\lau{This sentence does not compute}. However, to prevent various ways of misusing this basic scheme,
-we put local capabilities to work and take a number of not-always-obvious safety
-measures. In the next paragraphs, we will explain the issues to be considered in
-all the relevant situations: when (1) starting our program, (2) Returning to the
-adversary, (3) invoking the
-adversary, (4) returning from the adversary, (5) invoking an adversary callback
-and (6) having a callback invoked by the adversary.
+standard, C stack and register-passed return pointers, much like a standard C
+ABI\lau{This sentence does not compute}. However, to prevent various ways of
+misusing this basic scheme, we put local capabilities to work and take a number
+of not-always-obvious safety measures. In the next paragraphs, we will explain
+the issues to be considered in all the relevant situations: when (1) starting
+our program, (2) Returning to the adversary, (3) invoking the adversary, (4)
+returning from the adversary, (5) invoking an adversary callback and (6) having
+a callback invoked by the adversary.
 

### 2017-02-13 csf-2017/paper.tex
Removed outline section and added mention of sections in relevant contributions

@@
 \begin{itemize}
-\item We formalise a simple but representative capability machine featuring
+\item In Section~\ref{sec:capab-mach-with}, we formalise a simple but representative capability machine featuring
   local capabilities much like CHERI's, and define its operational semantics.
 \item We define a novel stack management discipline for enforcing control-flow
-  integrity and encapsulation of local state on the stack. The discipline relies
+  integrity and encapsulation of local state on the stack in Section~\ref{sec:stack-and-return-pointer}. The discipline relies
   solely on local capabilities (used to represent stack and return pointers) and
@@
   Dreyer et al.'s, except that they turn up in quite a different place
-  in the logical relation.
-\item We demonstrate the power of our results by applying them to challenging
+  in the logical relation. The logical relation and theorem is presented in Section~\ref{sec:logical-relation}.
+\item We demonstrate the power of our results in Section~\ref{sec:examples} by applying them to challenging
   examples, specifically constructed to demonstrate local state encapsulation
@@
 \end{itemize}
+\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution.
+Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}
 
@@
 % \end{itemize}
-
-\subsection{Outline}
-\todo[inline]{TODO}
 
@@
 The condition for the region to be address-stratified is there for
-technical reasons.\lau{Dominique, you probably disagree with this, so
-  please feel free to add something.}
+technical reasons.
 
@@
 \section{Examples}
+\label{sec:examples}
 In this section, we demonstrate how our formal model of capability
@@
 
-\lau{Add that the difficulties of the high-level version is also
-  present here nad solved in the same way?}
 \begin{figure}[t]
@@
 \end{figure}
+\begin{comment}
 \begin{itemize}
@@
 \end{itemize}
-
+\end{comment}
 \section{Discussion, future work}
+\label{sec:discussion}
 \begin{itemize}
@@
 \section{Related work}
+\label{sec:related-work}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2017-02-13 csf-2017/paper.tex
Took care of all the todos. Added two new ones.

@@
 and a register file, or a failed or halted state, where the latter keeps hold
-of the final state of memory. We also define memory segments $\ms$:
-parts of memory represented as partial functions mapping addresses to words.
-\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
+of the final state of memory. 
+% We also define memory segments $\ms$:
+% parts of memory represented as partial functions mapping addresses to words.
+%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
 
@@
 instructions. The instructions and data are opaque to an adversary holding $c$
-but when $c$ jumped to, the data becomes available to the closure's code through
+but when $c$ is jumped to, the data becomes available to the closure's code through
 the updated version of $c$ that is now in the $\pcreg$\footnote{Note that the
-  capability in the $\pcreg$-register can be used to load from.}. \lau{As we
-  talked about what is described here, is not how we make closures in the TR, so
-  we should consider whether we should refrain from using that word here. }
+  capability in the $\pcreg$-register can be used to load from.}.
+% \lau{As we
+%   talked about what is described here, is not how we make closures in the TR, so
+%   we should consider whether we should refrain from using that word here. }
 
@@
 
-Explain our model of linking
-\dominique{not sure what we want to say about it here...}
-
-Explain malloc
-\dominique{not sure what we want to say about it here either...}
+In order to have a reasonably realistic system, we have created a
+simple model of linking, so programs can have access to other programs
+through a linking table. We also assume malloc to be part of the
+trusted computing base satisfying a certain specification. Malloc and
+the model linking is described to the degree necessary, in the next
+section, but we refer to the technical appendix~\ref{XXX} for further
+details.
 
@@
 standard, C stack and register-passed return pointers, much like a standard C
-ABI\lau{This sentence does not compute}. However, to prevent various ways of
+ABI. However, to prevent various ways of
 misusing this basic scheme, we put local capabilities to work and take a number
@@
 
-\dominique{Add a summary table somewhere?}
+% \dominique{Add a summary table somewhere?}
+% \lau{It seems like we want have the space for this}.
 
@@
 return pointer, invoke a function in our code, wait for us to invoke him again
-and then jump to the stored, old return pointer rather than the new one he
-received upon the second invocation. \lau{If we need space, then maybe one of the examples is sufficient.}
-\lau{In the examples, we refer to the local return pointer as
-  ``protected return pointer''. We should either introduce that here
-  or use a different formulation in the examples. }
+and then jump to the stored, old return pointer rather than the new one they
+received upon the second invocation.
+% \lau{If we need space, then maybe one of the examples is sufficient.}
+% \lau{Both of these are good to have their as the two last examples of Section~\ref{sec:examples} use them.}
+%
+% \lau{In the examples, we refer to the local return pointer as
+%   ``protected return pointer''. We should either introduce that here
+%   or use a different formulation in the examples. }
+% \lau{It seemed did not seem to add anything to adopt the terminlogy - we
+%   often highlight that the return pointer is local which in most cases
+%   is so important that we do not want to leave it out.}
 
@@
 that it is safe to use.  Specifically, we check that it is indeed a $\rwlx$
-pointer and that its current address points just below its start address.
-Without these checks, err...
-\dominique{is this needed for security or just for convenience?}
-\lau{Say we are called with a stack capability that is global and $\rwx$ as well as an integer as a non-local return pointer (this is fair game as the adversary has to follow no rules).
-Now you push the l(a)unch codes to what you think is the stack and \texttt{scall} the callback.
-At this point \texttt{scall} does its thing, but because it never attempts to store anything local it will succeed and jump to the adversary!
-The adversary can now look at what you thought was the stack and eat your lunch or fire the missiles as they prefer.
-It is not sufficient to require the stack to be just local because there may be a global equivalent.
-We know this is not the case for the rwlx capability because we have a system wide assumption that all rwlx capabilities are local.
-Finally the adjustment of the pointer in \texttt{prepstck} (or whatever I called it) is a convenience/ space efficiency thing purely.}
-
-TODO:
-\begin{itemize}
-\item Reusable macro-instructions for stack and return pointer management
-\begin{itemize}
-\item taking care of the stuff above
-\end{itemize}
-\end{itemize}
-
-\lars{maybe we can present the examples informally before the logical relation and then return to them after the logical relation (to present the formal properties) ?}
+capability and make sure its current address points just below its start address.
+Without this check, an adversary could potentially get control over
+our local stack frame during a callback by passing us a local $\rwx$
+capability to a global data structure (i.e., there exists a global
+capability for it) as the stack capability and a global
+return capability. If our local state contains no local capabilities,
+then following our calling convention, the callback would not fail and
+the adversary could use a stored capability for the global data
+structure to access our local state. 
+To prevent this from happening, we need to make sure the stack
+capability is local and there exists no global capability for that
+memory segment. By making sure the capability has permission $\rwlx$,
+we rely on the system wide assumption that there are only local
+$\rwlx$ capabilities.
+
+\textbf{Reusable macro instructions} With the high-level approach in
+mind, we define the following macro instructions: 
+\texttt{fetch $r$ $\var{name}$} fetches the capability related to
+$\var{name}$ from the linking table and stores it in register
+$r$. 
+\texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
+macros that uses the stack assumes a stack capability in
+register $r_\stk$. 
+\texttt{prepstk $r$} prepares the stack by checking it has permission
+$\rwlx$ and making it conform to our stack discipline. 
+\texttt{scall  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}
+Jumps to the capability in register $r$ in a safe manner. That is it
+pushes local state, stack capability and in code return capability to
+the stack, sets up the necessary on stack return code and creates a
+return capability, restricts stack capability, clears the unused part
+of the stack, clears the necessary registers and jumpsto $r$. Upon
+return, the private state is restored.
+\texttt{mclear $r$} clears all the memory the capability in register
+$r$ has authority over.
+\texttt{rclear $\var{regSet}$} clears all the registers in $\var{regSet}$
+\texttt{reqglob $r$} checks whether the word in register $r$ is a
+global capability.
+\texttt{crtcls $\overline{(x_i,r_i)}$ $r$} allocates a closure pair
+where the capability in register $r$ is used to point to the program
+and a new environment is allocated where the contents of $\overline{r_i}$ is
+stored. In the program referred to by $r$, an implicit fetch
+happens when an instruction refers to $x_i$.
+
+In Section~\ref{sec:examples}, we use the above macros and some
+more. The technical appendix~\ref{XXX} contains detailed descriptions
+of all the macros.
+
+%\lars{maybe we can present the examples informally before the logical
+%relation and then return to them after the logical relation (to
+%present the formal properties) ?}
+% \lau{not enough space to do this.}
 
@@
 \subsection{Worlds}
-\todo[inline]{Find a good place to describe monotonicity of
-  interpretation function.}
+%\todo[inline]{Find a good place to describe monotonicity of
+%interpretation function.}
+% \lau{this may not be an important enough point to include it.}
 \begin{figure}[htbp]
@@
 $g$ is global.
-\todo[inline]{Do we want to mention the standard regions here? (They
-  are not mentioned anywhere)}
+The read condition uses one of two standard regions found
+in~\ref{fig:standard-regions}. $\iota^\pwl_{\start,\addrend}$ requires
+all the words in the range $[\start,\addrend]$ to be capability
+safe. $\iota^\nwl_{\start,\addrend}$ requires the words to be
+non-local and capability safe.
 
@@
 The interested reader can find all the technical details in the
-techincal report~\cite{XXX}.
+techincal appendix~\cite{XXX}.
 
@@
 state and uses an activation-record based calling convention
-(described in the technical report) to run the adversarial code.
+(described in the technical appendix) to run the adversarial code.
 
@@
 the assertion in the middle to succeed, they need to be
-well-bracketed. If the adversary is able to store the protected return
+well-bracketed. If the adversary is able to store the return
 pointer from the first call, so it is available for the second call, then it can be used
 in a context where \texttt{f3} has $2$ on top of its stack which will cause the
-assertion to fail. The protected return pointer is passed as a local
+assertion to fail. The return pointer is passed as a local
 capability, so the only means the adversary has to store it is the
@@
 so the adversary can only store it on the stack. This means that after
-the adversary jumps to the protected return pointer, \texttt{f3} can
+the adversary jumps to the return pointer, \texttt{f3} can
 revoke it (by erasing it which happens as part of the
 \texttt{scall}). In other words, the adversary cannot store the
-protected return pointer for the second call, and it is possible to
+return pointer for the second call, and it is possible to
 prove that the assertion never fails.
@@
 f4:  reqglob $r_1$
-     prepstack $r_\stk$
+     prepstk $r_\stk$
      store $x$ 0
@@
 an adversary can get to the first callback by calling the closure one
-more time. If the adversary can save the protected return pointer in
+more time. If the adversary can save the return pointer in
 the second callback and use it in the first, then the assertion will
-fail because $x$ is $0$ before the first callback. The protected return
+fail because $x$ is $0$ before the first callback. The return
 pointer cannot be transfered between the callbacks, so this does not

### 2017-02-13 csf-2017/paper.tex
Merge a discussion by Lars into a related work discussion that I accidentally rewrote without checking what was already there...

@@
 
-\subsection{Discussion}
-%/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
-%\todo[inline]{Add discussion/comparison to other LRs}
-% Comparisson of LR to other work:
-% - Single othorgonal closure
-%   o all capabilities are possible continuations, not just the one we
-%   intend to be it.
-% - No termination guarantees in observation relation
-% - Private/public transitions used like in Dreyer-Neis-Birkedal, but
-%   also to  handle the difference in what local and global
-%   capabilities can handle.
-
-\todo[inline]{Maybe move this subsection to related work}
-
-As already mentioned, our the state transition systems used in our
-logical relation are inspired by earlier work on logical relations for
-high-level
-languages~\cite{Ahmed:popl09, Dreyer:jfp12,Devriese:2016ObjCap}.
-The idea of public and private transitions comes from 
-Dreyer et. al.~\cite{Dreyer:jfp12}, who gave a logical relations model
-of a high-level lambda calculus. In \emph{loc. cit.}
-public transitions were used to 
-capture the \emph{extensional} behavior of a program expression
-(and a context), whereas private transitions were used to capture
-the internal behaviour of a computation using local state.
-Technically, the logical relation of Dreyer et. al.\ was
-formulated using biorthogonal closure, and the future world
-relation for continuations was required to be a public future world
-relation. In other words, the choice between private or public future
-rested on the ``structure'' of the programming language. 
-In contrast, 
-for our low-level, unstructured, assembly language, we
-we instead decide on whether to use a private or public future 
-world depending on whether the capability is global or local. 
-Another interesting technical difference is that in our low-level 
-language we do not have a fixed calling convention and any register
-can serve as a continuation, and therefore we do not use biorthogonal
-closure but just model return pointers as continuations using a single
-orthogonal closure.
-
-\lars{don't know yet whether we need something like the following paragraph:}
-Even though 
-Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also
-considered a logical
-relation involving an assembly language, they still used
-biorthogonal closure but that was because their logical relation
-was a relation between a high-level language and an assembly
-language,  and thus the assembly language in the relation essentially
-inherited structure from the high-level language.
-
 \section{Examples}
@@
 \end{comment}
+
 \section{Discussion, future work}
 \label{sec:discussion}
+
 \begin{itemize}
@@
   \end{itemize}
-\item Relation to local parameters in Scala, Algol/Pascal second-class function parameters?
+\item Relation to local parameters in Scala, Algol/Pascal second-class function
+  parameters? 
 \item Stack clearing realistic?
@@
 \end{itemize}
+
+It is worth mentioning that local capabilities are quite similar to features
+that have been proposed for high-level languages. Specifically,
+\citet{osvald_gentrification_2016}'s second-class or local values in Scala are
+remarkably similar and we expect they will support similar reasoning patterns.
 
@@
 fundamental theorem is similar to theirs and can be seen as a formulation of
-capability-safety of the capability-machine. Because we are not interested in an
-externally observable notion of side-effects (like console output or memory
-access traces), we do not require their notion of effect parametricity.
+capability-safety of the capability-machine. Because we are not interested in
+externally observable side-effects (like console output or memory access
+traces), we do not require their notion of effect parametricity.
 
@@
 condition on the assembly code are quite loose (for example, they allow
-hand-written and self-modifying assmebly code), but they do assume code that
-is non-adversarial and respects the ABI of the high-level language (for example,
-the code does not store return pointers and invoke them after their intended
-lifetime). As a result, they can treat return pointers as evaluation contexts in
-a lambda calculus and use biorthogonality~\cite{pitts_operational_1998}. This
-means that they define the logical relation to accept evaluation contexts if
-they produce related observations for related values and to accept terms if they
-produce related observations when plugged into related evaluation contexts. In
-our setting, there is no guarantee that adversarial code will treat return
-pointers as intended, so, as described in Section~\ref{sec:logical-relation}, we
-define our logical relation in a (arguably natural but still) novel way, using a single
-orthogonal closure, i.e. accept terms if they produce related observations when
-given related arguments, where those arguments include the return pointer.
+hand-written and self-modifying assembly code), but they do assume code that is
+non-adversarial and respects the calling convention of the high-level language
+(for example, the code does not store return pointers and invoke them after
+their intended lifetime). As a result, they can treat return pointers as
+evaluation contexts (continuations) in a lambda calculus and use biorthogonality
+to define the LR~\cite{pitts_operational_1998}. In our setting, there is no
+guarantee that potentially adversarial code will respect the calling convention,
+so, as described in Section~\ref{sec:logical-relation}, we do not use
+biorthogonal closure but just model return pointers as continuations using a
+single orthogonal closure.
 
@@
 nature of local capabilities. The main insight is that the temporary, revokable
-nature of these capabilities can be seen as a generalisation of the special
-status of evaluation contexts in lambda calculi. Closely related to this is a
-study by \citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and
-control effects on local relational reasoning. DNB formally capture this special
-status of evaluation contexts using Kripke worlds with public and private future
-world relations. Our public and private worlds are inspired by theirs and they
-are at times remarkably similar. For example, the state-transitition system in
-the proof of our assembly version of the awkward example, is exactly the same as
-the one used by DNB for the original awkward example. Nevertheless, there are
-also remarkable differences: in DNB's logical relation, public future worlds
-only turn up in the continuation relation (which we do not have, see above).
-This is different from our value relation which uses public and private future
-worlds depending on the locality of the capability at hand. Additionally, our
-worlds are a bit more complex because, to allow store-local capabilities, they
-can have temporary regions that are only monotonous w.r.t. public future worlds,
-while DNB's worlds are entirely permanent.
-
-As a side note: our model of local/global capabilities may also be applicable
-for reasoning about related ideas in higher-level languages. Specifically,
-\citet{osvald_gentrification_2016}'s second-class or local values in Scala are
-remarkably similar to the local capabilities we study and we expect they will
-support similar reasoning patterns.
+nature of these capabilities can be seen as a generalisation of the
+syntactically special, unstoreable status of evaluation contexts in lambda
+calculi (without control effects). Closely related to this is a study by
+\citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and control
+effects on local relational reasoning. DNB formally capture the special status
+of evaluation contexts using Kripke worlds with public and private future world
+relations. Our future world relations are inspired by theirs and they are at
+times remarkably similar. For example, the state-transitition system in the
+proof of our assembly version of the awkward example, is exactly the same as the
+one used by DNB for the original awkward example. However, in our logical
+relation, public and private future worlds turn up in an entirely different
+place in the logical relation: rather than using public future worlds for the
+special syntactic category of evaluation contexts, they are used in the value
+relation depending on the depending on the locality of the capability at hand.
+Additionally, our worlds are a bit more complex because, to allow store-local
+capabilities, they can have temporary regions that are only monotonous w.r.t.
+public future worlds, while DNB's worlds are entirely permanent.
+
+%/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
+%\todo[inline]{Add discussion/comparison to other LRs}
+% Comparisson of LR to other work:
+% - Single othorgonal closure
+%   o all capabilities are possible continuations, not just the one we
+%   intend to be it.
+% - No termination guarantees in observation relation
+% - Private/public transitions used like in Dreyer-Neis-Birkedal, but
+%   also to  handle the difference in what local and global
+%   capabilities can handle.
+
+% As already mentioned, our the state transition systems used in our
+% logical relation are inspired by earlier work on logical relations for
+% high-level
+% languages~\cite{Ahmed:popl09, Dreyer:jfp12,Devriese:2016ObjCap}.
+% The idea of public and private transitions comes from 
+% Dreyer et. al.~\cite{Dreyer:jfp12}, who gave a logical relations model
+% of a high-level lambda calculus. In \emph{loc. cit.}
+% public transitions were used to 
+% capture the \emph{extensional} behavior of a program expression
+% (and a context), whereas private transitions were used to capture
+% the internal behaviour of a computation using local state.
+% Technically, the logical relation of Dreyer et. al.\ was
+% formulated using biorthogonal closure, and the future world
+% relation for continuations was required to be a public future world
+% relation. In other words, the choice between private or public future
+% rested on the ``structure'' of the programming language. 
+% In contrast, 
+% for our low-level, unstructured, assembly language, we
+% we instead decide on whether to use a private or public future 
+% world depending on whether the capability is global or local. 
+% Another interesting technical difference is that in our low-level 
+% language we do not have a fixed calling convention and any register
+% can serve as a continuation, and therefore we do not use biorthogonal
+% closure but just model return pointers as continuations using a single
+% orthogonal closure.
+
+% \lars{don't know yet whether we need something like the following paragraph:}
+% Even though 
+% Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also
+% considered a logical
+% relation involving an assembly language, they still used
+% biorthogonal closure but that was because their logical relation
+% was a relation between a high-level language and an assembly
+% language,  and thus the assembly language in the relation essentially
+% inherited structure from the high-level language.
 

### 2017-02-13 csf-2017/paper.tex
Added a small disclaimer that in the paper there really is no distinction between pointer and capability. Made sure that we use the same terms (stack pointer and return pointer) everywhere. Removed a couple of unnecessary hyphens.

@@
   inspired by one used in \cite{Devriese:2016ObjCap} for
-  capability-safety in a JavaScript-like lambda calculus, but has two
+  capability safety in a JavaScript-like lambda calculus, but has two
   novel technical characteristics. The natural CPS style of assembly
@@
   function pointers. The examples demonstrate both the power of our formulation
-  of capability-safety and our stack management discipline.
+  of capability safety and our stack management discipline.
 \end{itemize}
@@
 
-We define the syntax of our capability machine in \figurename~\ref{fig:syntax}. We
-assume an infinite set of addresses $\Addrs$ and define machine words as either
-integers or capabilities of the form $((\perm,\gl),\start,\addrend,\addr)$. Such
-a capability represents the authority to execute permissions $\perm$ on the
-memory array $[\start,\addrend]$, together with a current address $\addr$ and a
-tag $\gl$ which indicates whether the capability is global or local. The
-available permissions and their ordering are depicted in
-\figurename~\ref{fig:perm-hier}: the permissions include null permission ($\noperm$),
-readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and
-read/write/execute ($\rwx$) permissions. Additionally, there are three special
-read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter
+We define the syntax of our capability machine in
+\figurename~\ref{fig:syntax}. We assume an infinite set of addresses
+$\Addrs$ and define machine words as either integers or capabilities
+of the form $((\perm,\gl),\start,\addrend,\addr)$. Such a capability
+represents the authority to execute permissions $\perm$ on the memory
+array $[\start,\addrend]$, together with a current address $\addr$ and
+a tag $\gl$ which indicates whether the capability is global or
+local. The capability machine has no notion of pointers by itself, so
+we refer to capabilities as pointers and capabilities
+interchangeably. The available permissions and their ordering are
+depicted in \figurename~\ref{fig:perm-hier}: the permissions include
+null permission ($\noperm$), readonly ($\readonly$), read/write
+($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$)
+permissions. Additionally, there are three special read/write-local
+($\readwritel$), read/write-local/execute ($\rwlx$) and enter
 ($\entry$) permissions which we will explain below.
@@
 capability to a global data structure (i.e., there exists a global
-capability for it) as the stack capability and a global
-return capability. If our local state contains no local capabilities,
+capability for it) as the stack pointer and a global
+return pointer. If our local state contains no local capabilities,
 then following our calling convention, the callback would not fail and
@@
 \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
-macros that uses the stack assumes a stack capability in
+macros that uses the stack assumes a stack pointer in
 register $r_\stk$. 
@@
 Jumps to the capability in register $r$ in a safe manner. That is it
-pushes local state, stack capability and in code return capability to
+pushes local state, stack pointer and in code return pointer to
 the stack, sets up the necessary on stack return code and creates a
-return capability, restricts stack capability, clears the unused part
+return pointer, restricts stack pointer, clears the unused part
 of the stack, clears the necessary registers and jumpsto $r$. Upon
@@
 whether the local state is unchanged. They differ in the way they do
-this. Program \texttt{f1} assumes it has access to a stack capability
+this. Program \texttt{f1} assumes it has access to a stack pointer
 and uses the stack to store its local state. It also uses the
@@
 executing the program (see Lemmas~\ref{lem:correctness-f1}
-and~\ref{lem:correctness-f2} below).  \lars{shouldn't we say already
-here that the adversary is a global capability and that is why,
-intuitively, the local state is preserved ?} \lau{Due to the ``order
-of control'', \texttt{f1} runs first. The lemma is quite restrictive
-on the adversary in the sense that it only has access to malloc (and
-whatever we give et access to). Specifically, it does not have access
-to the stack capability and because of that we don't need to check it
-first. We could write that the lemmas are limited in their assumptions
-to make them simpler and a more complicated example is coming up.}
+and~\ref{lem:correctness-f2} below).
+% \lars{shouldn't we say already
+% here that the adversary is a global capability and that is why,
+% intuitively, the local state is preserved ?} \lau{Due to the ``order
+% of control'', \texttt{f1} runs first. The lemma is quite restrictive
+% on the adversary in the sense that it only has access to malloc (and
+% whatever we give et access to). Specifically, it does not have access
+% to the stack pointer and because of that we don't need to check it
+% first. We could write that the lemmas are limited in their assumptions
+% to make them simpler and a more complicated example is coming up.}
 The two examples also illustrate the versatility of the logical
@@
 \]
-  where each of the capability has an appropriate range of authority and pointer and
+  where each of the capabilities have an appropriate range of authority and pointer furthermore
   \begin{itemize}
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} but assuming no stack capability
+  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer
   if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
@@
 capability, so the only means the adversary has to store it is the
-stack capability. The stack capability itself is a local capability,
+stack pointer. The stack pointer itself is a local capability,
 so the adversary can only store it on the stack. This means that after
@@
 callback is not a piece of code on the stack that restores the
-adversaries original stack capability breaking the encapsulation of
+adversaries original stack pointer breaking the encapsulation of
 our local stack frame. In the proof of Lemma~\ref{lem:correctness-g1},
@@
 closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
-capability-safety in a high-level JavaScript-like lambda calculus. Our
+capability safety in a high-level JavaScript-like lambda calculus. Our
 fundamental theorem is similar to theirs and can be seen as a formulation of
-capability-safety of the capability-machine. Because we are not interested in
+capability safety of the capability machine. Because we are not interested in
 externally observable side-effects (like console output or memory access

### 2017-02-13 csf-2017/paper.tex
Added a couple of essential details to g1 lemma

@@
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
@@
 return pointer for the second call, and it is possible to
-prove that the assertion never fails.
-\begin{figure}[htbp]
+prove that the assertion never fails. The correctness lemma for \texttt{f3} is similar to the two previous lemmas:
+\begin{figure}[t]
   \centering
@@
 \end{figure}
-\lau{Add something about the correctness lemma? Not sure if I have
-  anything important to say.}
+
 \begin{lemma}
@@
   \label{lem:correctness-g1}
-  Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} 
+  Let
+\[
+    \begin{array}{r@{}lr@{}l}
+    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
+    \end{array}
+\]
+  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1} 
   if

### 2017-02-14 csf-2017/paper.tex
Some math changes (please review!) - changed '\start' to be 'b' and \addrend to be 'e'. Makes the figure fit in their columns, but may affect readability.

@@
 \newcommand{\word}{\var{w}}
-\newcommand{\start}{\var{base}}
-\newcommand{\addrend}{\var{end}}
+\newcommand{\start}{\var{b}}
+\newcommand{\addrend}{\var{e}}
 \newcommand{\pwlv}{\var{pwl}}
@@
 $\Addrs$ and define machine words as either integers or capabilities
-of the form $((\perm,\gl),\start,\addrend,\addr)$. Such a capability
+of the form $((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability
 represents the authority to execute permissions $\perm$ on the memory
-array $[\start,\addrend]$, together with a current address $\addr$ and
+array $[\var{base},\var{end}]$, together with a current address $\addr$ and
 a tag $\gl$ which indicates whether the capability is global or
@@
   &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \; \middle| \;
+  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
     \begin{aligned}
@@
   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \\
-    &\quad\left\{\npair{\hs} \; \middle| \;
-    \begin{multlined}
-      n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \forall \addr \in [\start,\addrend] \ldotp \\
-       {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }} \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-    \end{multlined} \right\}
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
+    \begin{aligned}
+      &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
+      &\forall \addr \in [\start,\addrend] \ldotp \\
+      & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
+      & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
+     \end{aligned}
+     \right\}
 \end{align*}
@@
     \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp  W(r) \text{\footnotesize{ is address-stratified and}} \\
+      & \exists r \in \var{localityReg}(g,W) \ldotp \\
+      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
       & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}

### 2017-02-14 csf-2017/paper.tex
put terminology section in comment

@@
 
-\section*{Terminology ?}
-
-\emph{capability safe}: 
-Our unary logical relation is a semantic \emph{definition}
-of which values for the
-pc / register-file / word are capability safe.
-
-\emph{adversary}: It doesn't sound ``sexy'' to constantly say ``an untrusted
-piece of code'' should we adopt the convention that adversary refers
-to an unstrusted piece of code?
-
-\emph{memory}  we don't make a distinction between memory and memory
-segments in the text  (to do so properly, we would need to explain why
-assumptions
-are reasonable, given that a real machine has finite memory, but we
-need infinite domain for malloc spec, etc.)
+% \section*{Terminology ?}
+
+% \emph{capability safe}: 
+% Our unary logical relation is a semantic \emph{definition}
+% of which values for the
+% pc / register-file / word are capability safe.
+
+% \emph{adversary}: It doesn't sound ``sexy'' to constantly say ``an untrusted
+% piece of code'' should we adopt the convention that adversary refers
+% to an unstrusted piece of code?
+
+% \emph{memory}  we don't make a distinction between memory and memory
+% segments in the text  (to do so properly, we would need to explain why
+% assumptions
+% are reasonable, given that a real machine has finite memory, but we
+% need infinite domain for malloc spec, etc.)
 

### 2017-02-14 csf-2017/paper.tex
Some slight changes in the text about macro-instructions.

@@
 
-\textbf{Reusable macro instructions} With the high-level approach in
-mind, we define the following macro instructions: 
-\texttt{fetch $r$ $\var{name}$} fetches the capability related to
-$\var{name}$ from the linking table and stores it in register
-$r$. 
-\texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
-macros that uses the stack assumes a stack pointer in
-register $r_\stk$. 
-\texttt{prepstk $r$} prepares the stack by checking it has permission
-$\rwlx$ and making it conform to our stack discipline. 
-\texttt{scall  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}
-Jumps to the capability in register $r$ in a safe manner. That is it
-pushes local state, stack pointer and in code return pointer to
-the stack, sets up the necessary on stack return code and creates a
-return pointer, restricts stack pointer, clears the unused part
-of the stack, clears the necessary registers and jumpsto $r$. Upon
-return, the private state is restored.
-\texttt{mclear $r$} clears all the memory the capability in register
-$r$ has authority over.
-\texttt{rclear $\var{regSet}$} clears all the registers in $\var{regSet}$
-\texttt{reqglob $r$} checks whether the word in register $r$ is a
-global capability.
-\texttt{crtcls $\overline{(x_i,r_i)}$ $r$} allocates a closure pair
-where the capability in register $r$ is used to point to the program
-and a new environment is allocated where the contents of $\overline{r_i}$ is
-stored. In the program referred to by $r$, an implicit fetch
+\textbf{Reusable macro instructions} With the high-level approach in mind, we
+define the following macro instructions: \texttt{fetch $r$ $\var{name}$} fetches
+the capability related to $\var{name}$ from the linking table and stores it in
+register $r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
+macros that uses the stack assumes a stack pointer in register $r_\stk$.
+\texttt{prepstk $r$} is used when a callback is invoked by the adversary and
+prepares the received stack pointer by checking that it has permission $\rwlx$
+and making it conform to our stack discipline. \texttt{scall
+  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps to the
+capability in register $r$ in the manner described above. That is, it pushes local state (the
+contents of registers $\overline{r_{\var{priv}}}$), stack pointer and return
+pointer to the stack, sets up the necessary on stack return code and creates a
+return pointer, restricts stack pointer, clears the unused part of the stack,
+clears the necessary registers and jumps to $r$. Upon return, the private state
+is restored. \texttt{mclear $r$} clears all the memory the capability in
+register $r$ has authority over. \texttt{rclear $\var{regSet}$} clears all the
+registers in $\var{regSet}$ \texttt{reqglob $r$} checks whether the word in
+register $r$ is a global capability. \texttt{crtcls $\overline{(x_i,r_i)}$ $r$}
+allocates a closure pair where the capability in register $r$ is used to point
+to the program and a new environment is allocated where the contents of
+$\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch
 happens when an instruction refers to $x_i$.
@@
 % \lau{this may not be an important enough point to include it.}
-\begin{figure}[htbp]
+\begin{figure}
   \centering
@@
 
-We believe local capabilities
-
 Other papers that enforce well-bracketed control flow at low level (using a

### 2017-02-14 csf-2017/paper.tex
Some details about closures.

@@
 access to, as well as some prelude instructions that load the data into
-registers and jumps to the right code. We can then provide the adversary with an
+registers and jumps to the right code. Note that this implies that no local
+capabilities can be stored as part of a closure. We can then provide the adversary with an
 enter-capability covering the allocated block and pointing to the contained
@@
 allocates a closure pair where the capability in register $r$ is used to point
-to the program and a new environment is allocated where the contents of
+to the program and a new environment is allocated (using malloc) where the contents of
 $\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch

### 2017-02-14 csf-2017/paper.tex
First attempt at an abstract

@@
 \begin{abstract}  
-  abc
+  Capability machines provide security guarantees at machine level
+  which makes them an interesting target for compilation schemes that aim to
+  provably enforce control-flow integrity, encapsulation of local
+  state, and such properties.
+  We provide a formalisation of a representative capability machine
+  and propose a novel stack management discipline for enforcing
+  control-flow integrity and encapsulation of local state on it. To prove
+  these properties, we provide a logical relation that semantically
+  captures capability safety. The capability safety defined by the
+  logical relation goes beyond the
+  proposed stack management discipline and can be used to reason about
+  any capability safe program.
 \end{abstract}

### 2017-02-14 csf-2017/paper.tex
removed comment

@@
 \label{sec:stack-and-return-pointer}
-\lau{Do we want to use the term ``pointer'' at any point? I have also used it, but I guess it might create some confusing to whether we have traditional pointer as well as capabilities?}
+%\lau{Do we want to use the term ``pointer'' at any point? I have also used it, but I guess it might create some confusing to whether we have traditional pointer as well as capabilities?}
 One of the contributions in this paper is a demonstration that local

### 2017-02-14 csf-2017/paper.tex
added sentence about stack being abstraction on memory
Some notes and small changes to the LR section.

--- 2017-02-14 csf-2017/paper.tex
@@
 machine and prove properties such as well-bracketedness and
-encapsulation about them, we define a unary logical relation.
+encapsulation, we define a unary logical relation.
 % Intuition
-The logical relation is a semantic model of all the words on the
-capability machine that are definitely safe to give to an
-adversary. Whether a capability is safe to give to an adversary
-depends heavily on the contents of the memory since instructions as well
-as capabilities may reside in memory. Hence we index the logical
-relation over worlds, which are used to model the memory, and our logical
-relation is thus an example of what is often referred to as a Kripke
-logical relation. A world consists of a finite set of regions, and each region
-can be understood as a protocol describing the possible evolution of
-part of the memory. Memory changes during an
-execution since new memory can be allocated and part of the existing
-memory can change. We use two future world relations to capture
-such changes to the memory in our model.
+It is a semantic model of all the words on the capability machine that are
+definitely safe to give to an adversary. Whether a capability is safe to give to
+an adversary depends heavily on the contents of the memory since instructions as
+well as capabilities may reside in memory. Hence we index the logical relation
+over worlds, which are used to model the memory (making our logical relation a
+Kripke logical relation). A world consists of a finite set of regions, and each
+region can be understood as a protocol describing the possible evolution of part
+of the memory. Future evolutions of worlds (e.g. new protocols established over
+memory received from malloc, or steps in a protocol already established for a
+part of memory) are modeled using a future world relation. In fact, as we will
+explain below, there are two such relations to accomodate reasoning about local
+capabilities.
 %
@@
 \subsection{Worlds}
+\dominique{Perhaps we could first try to convey the intuition of a permanent
+  world with a temporary overlay, with global capabilities ignoring the overlay.
+  Perhaps we could add a depiction of the world to Figure 5?}
+
 %\todo[inline]{Find a good place to describe monotonicity of
@@
 public and private transitions, which describes the protocol governing
-the evolution of the memory described by the region, similarly to what
+the memory described by the region, similarly to what
 has been used in logical relations for high-level
@@
 regions can be revoked at certain points during execution,
-corresponding to the points when local capabilities can be
+corresponding to the points, on the condition that we clear local capabilities can be
 revoked. This means that local capabilities safely can depend on the
@@
 
-For technical reasons (monotonicity of the logical relation), a
-temporary region is not removed in the model when it is
-revoked. Instead it is turned into a new kind of region, namely a
-revoked region. A revoked region contains no state transition
-system and puts no requirements on the memory it models. It
-simply serves as a mask for a temporary region that have been revoked.
-This way of masking a region goes back to earlier work of
-Ahmed~\cite{Ahmed2004semantics} and was also used in~\cite{Thamsborg:2011:KLR:2034773.2034831}.
+For technical reasons, a temporary region is not removed in the model when it is
+revoked. Instead it is turned into a new kind of region, namely a revoked
+region. A revoked region contains no state transition system and puts no
+requirements on the memory it models. It simply serves as a mask for a
+temporary region that has been revoked. This way of masking a region goes back
+to earlier work of \citet{Ahmed2004semantics} and was also used
+by \citet{Thamsborg:2011:KLR:2034773.2034831}.
 
@@
 
-The worlds end up being defined recursively: Worlds describe memory,
-which can contain capabilities, whose safety depends on worlds. 
-We use the method of Birkedal et
-al.~\cite{Birkedal:2011:SKM:1926385.1926401} 
-for constructing the worlds.
+\dominique{``end up being defined recursively'' does not sound very insightful}
+The worlds end up being defined recursively: Worlds describe memory, which can
+contain capabilities, whose safety depends on worlds. We use the method of
+\citet{Birkedal:2011:SKM:1926385.1926401} for constructing the worlds.
 
@@
 % recursive domain equation.
+\dominique{it is not explained what cofes are?}
 \begin{theorem}\label{thm:world-existence}
@@
 \end{mathpar}
-These rules express that, in the private future region relation, 
-we allow revocation of temporary regions. 
-As explained above, this can be done using a special revoked region,
-which does not describe any memory, but we also allow  
-any other region to serve as a mask, and thus any region $r$ can be a
-future region of a temporary region and of a revoked region.
-If a temporary region is not masked by a region, then it behaves like a
-permanent region. The permanent regions stay permanent and they are
-allowed to take a transition in the private part of the transition
-system.
-
-The \emph{public future} region relations satisfies the following properties:
+These rules express that, in the private future region relation, we allow
+revocation of temporary regions. As explained above, this can be done using a
+special revoked region, which does not describe any memory, but we also allow
+any other region to serve as a mask, and thus any region $r$ can be a future
+region of a temporary region and of a revoked region. On the other hand,
+permanent regions cannot be masked away. They are only allowed to take a
+transition in the private part of the transition system.
+
+The \emph{public future} region relation satisfies \dominique{is defined by?} the following properties:
 \begin{mathpar}
@@
 \end{mathpar}
-In the public future regions we do not permit revocation of temporary
-regions. Instead, temporary and permanent regions are only 
-allowed to take a transition in the public part of the transition system.
-In the public future region relation, we only allow revoked and
-temporary regions to mask a region. This means that the private future
-world relations allows us to reuse a region that has been revoked
+In the public future regions we do not permit revocation of temporary regions.
+Instead, temporary and permanent regions are only allowed to take a transition
+in the public part of the transition system. Revoked regions can either remain
+revoked or be masked away by a temporary region. This means that the public
+future world relations allows us to reuse a region that has been revoked
 earlier.
@@
 %subsection introduction
-The logical relation defines semantically what it means to be
-capability safe. In large parts, capability safety is defined by
-when a capability safely can be given to an untrusted piece of
-code. In the following paragraphs, we will go through the definitions
-of the logical relation and provide an intuition why they define
-capability safety.
-The definition of the logical relation is found in
-\figurename~\ref{fig:logrel}, \figurename~\ref{fig:standard-regions},
-and \figurename~\ref{fig:perm-conds}. Some definitions are omitted and
-explained when needed. The precise definitions can be found in the
-technical appendix~\cite{XXX}.
-
-The \emph{observation relation}, $\observations$, defines which
-configurations are capability safe. Roughly speaking,
-a configuration is capability safe if, whenever it halts, then
-the resulting memory satisfies a private future world. 
-Notice that if a configuration fails, e.g., if an unauthorized access 
-(e.g., a load of a memory location without the requisite read capability)
-is attempted, then the configuration is considered capability safe, which it should
-because then the capability machine has caught the attempted unauthorized access
-This is similar to the logical relation given for an untyped language
-in~\cite{Devriese:2016ObjCap}, but unlike logical relations for typed languages, which
-typically require that computations do not fail.
-
-The \emph{register-file relation}, $\stdrr$, specifies all the
-register-files that are capability safe. Capability safe
-register-files contain capability safe words (e.g.n, words in
-$\stdvr$) in all registers but $\pcreg$.
-
-The \emph{expression relation}, $\stder$, defines when a word is
-capability safe as a program counter. That is the case when the word can be plugged into a
-capability safe register file (i.e., a register file in $\stdrr$) and
-paired with a memory satisfying the world to become a capability safe
-configuration.
-
-The \emph{value relation}, $\stdvr$, defines which words are capability
-safe. A word is capability safe if it cannot be used to break the
-protocols defined by the world. Data is always capability safe because
-it provides no authority that can be used to break the
-invariants. Capabilities gives the authority to manipulate memory and
-potentially break memory protocols, so they need to satisfy certain
-conditions to be capability safe. We define a condition for each kind
-of permission a capability can have. These permission-based conditions
-can be found in \figurename~\ref{fig:perm-conds}.
-
-The condition for capabilities with read permission is
-$\readCond{}$. It ensures that a read capability can only be used to
-read capability safe words, by requiring that the region that governs
-the part of memory that the capability grants authority over enforces
-that all the words are in the value relation. Generally speaking,
-global capabilities should only depend on permanent regions. Therefore
-we use the function $\var{localityReg}(g,W)$, which projects out all
-active regions when $g$ is local, but only the permanent regions when
-$g$ is global.
-The read condition uses one of two standard regions found
-in~\ref{fig:standard-regions}. $\iota^\pwl_{\start,\addrend}$ requires
-all the words in the range $[\start,\addrend]$ to be capability
-safe. $\iota^\nwl_{\start,\addrend}$ requires the words to be
-non-local and capability safe.
-
-For a capability with a write permission, $\writeCond{}$ must be
-satisfied for the capability's range of authority. Generally speaking,
-write capabilities may be used to write any word, so the region that
-governs the part of memory the write capability has authority over
-must at least allow any capability safe word to be stored.
-This is of course not entirely true because of the distinction between
-write and permit-write-local capabilities. For write capabilities, the
-region must require the contents of memory to be non-local (and
-capability safe). For permit-write-local capabilities, the
-non-local requirement is omitted.
-The write condition uses $\var{localityReg}$ to pick an appropriate
-region in the same way the read condition does.
-The condition for the region to be address-stratified is there for
-technical reasons.
-
-The conditions $\entryCond{}$ and $\execCond{}$ are very similar and
-differ only slightly. Executable capabilities can be updated to point
-anywhere in memory, so they must be capability safe as a program
-counter (in the $\stder$-relation) no matter where they point in memory. In contrast, enter
-capabilities are opaque and must be used from the address they point
-to. They change permission when jumped to, so we require them to be
-capability safe as a program counter when the permission is changed to
-read/execute.
-In both cases, it must be safe to jump to the capability. This
-means that the capability must be capability safe as a
-program counter in a future world. If the capability is global,
-then it should only depend on permanent regions and therefore
-we require that it should be capability safe as a program counter in any \emph{private}
-future world (recall that temporary regions may be revoked in a private future
-world). On the other hand, a local capability should be capability
-safe in any \emph{public} future world because it may depend on temporary
-regions, so we cannot expect them to function correctly as program
-counters if temporary regions are revoked.
-
-The definition of the value relation in \figurename~\ref{fig:logrel}
-is defined in terms of the permission-based conditions.
+\dominique{I would try to motivate the LR defs by saying that the LR is
+  constructed to express the FTLR and the FTLR says that if a piece of code only
+  has access to values respecting the invariants in a world, then it will
+  respect them itself too.} The logical relation defines semantically what it
+means to be capability safe.\dominique{We have not introduced capability
+  safety..} In large parts, capability safety is defined by when a capability
+can be safely given to an untrusted piece of code. In the following paragraphs,
+we will go through the definitions of the logical relation and provide an
+intuition why they define capability safety. The definition of the logical
+relation is found in \figurename~\ref{fig:logrel},
+\figurename~\ref{fig:standard-regions}, and \figurename~\ref{fig:perm-conds}.
+Some definitions are omitted and explained when needed. The precise definitions
+can be found in the technical appendix~\cite{XXX}.
+
+The \emph{observation relation} $\observations$ defines which configurations
+are capability safe. Roughly speaking, a configuration is capability safe if,
+whenever it halts, then the resulting memory satisfies a private future world.
+Notice that failing is considered safe behavior and in fact, the machine will
+often resort to failing when, for example, an unauthorized
+access is attempted, like a load of a memory location through a
+capability without read permission. This is similar to the logical relation
+given for an untyped language in~\cite{Devriese:2016ObjCap}, but unlike logical
+relations for typed languages, which typically require that computations do not
+fail.
+
+The \emph{register-file relation} $\stdrr$ specifies that register-files are
+safe when they contain safe words (i.e., words in $\stdvr$) in all registers but
+$\pcreg$. The \emph{expression relation} $\stder$ defines when a word is safe
+to use as a program counter. That is the case when the word can be plugged into
+a capability safe register file (i.e., a register file in $\stdrr$) and paired
+with a memory satisfying the world to become a capability safe configuration.
+
+The \emph{value relation} $\stdvr$ defines when words are safe, i.e. when they
+cannot be used to break the protocols defined by the world. Non-capability data
+is always safe because it provides no authority. Capabilities give the authority
+to manipulate memory and potentially break memory protocols, so they need to
+satisfy certain conditions to be capability safe. In
+\figurename~\ref{fig:perm-conds}, we define such a condition for each kind of
+permission a capability can have (we will call these permission-based
+conditions).
+
+For capabilities with read permission, the $\readCond{}$ ensures that it can
+only be used to read capability safe words, by requiring that the region
+governing the relevant part of memory, guarantees that the memory contains only
+words in the value relation. Generally speaking, global capabilities should only
+depend on permanent regions. Therefore we use the function
+$\var{localityReg}(g,W)$, which projects out all active regions when $g$ is
+local, but only the permanent regions when $g$ is global. The read condition
+uses $\iota^\pwl_{\start,\addrend}$, a standard regions defined in
+Figure~\ref{fig:standard-regions}, which requires all the words in the range
+$[\start,\addrend]$ are safe.
+
+For a capability with write permission, $\writeCond{}$ must be satisfied for the
+capability's range of authority. Such a capability may be used to write any
+word, so the region governing the relevant part of memory must (at least) allow
+any capability safe word to be stored. However, this is not entirely true
+because of the distinction between write and permit-write-local capabilities.
+Only permit-write-local capabilities need to allow arbitrary safe values to be
+stored, while for write capabilities, it is sufficient to allow storing values
+that are both non-local and safe. The write condition uses $\var{localityReg}$
+to pick an appropriate region for its locality in the same way the read
+condition does. The condition for the region to be address-stratified is there
+for technical reasons.
+
+The conditions $\entryCond{}$ and $\execCond{}$ are very similar and differ only
+slightly. In both cases, we require that the capability can be safely jumped to.
+However, executable capabilities can be updated to point anywhere in memory, so
+they must be capability safe as a program counter (in the $\stder$-relation) no
+matter where they point to. In contrast, enter capabilities are opaque and can
+only be used to jump to the address they point to. They also change permission
+when jumped to, so we require them to be capability safe as a program counter
+when the permission is changed to read/execute. Because the capabilities are not
+necessarily invoked immediately, we need to require that this is the case in a
+future world, but it depends on the locality of the capability which future
+worlds we consider. If the capability is global, then it should only depend on
+permanent regions, so we require capability safety as a program counter in any
+\emph{private} future world (recall that temporary regions may be revoked in a
+private future world). For local capabilities, it suffices to be capability safe
+in \emph{public} future worlds, where temporary regions are still there.
 

### 2017-02-14 csf-2017/paper.tex

@@
 for which we receive a stack pointer in a register, $r_\stk$, that we reserve for this
-purpose. This stack pointer must be local and it must carry $\rwlx$ permission. It is crucial for
+purpose. We stress that the stack is not built-in, but merely an
+abstraction we put on a piece of the memorey. This stack pointer must be local and it must carry $\rwlx$ permission. It is crucial for
 our approach that the stack is the only part of memory for which there will ever

### 2017-02-14 csf-2017/paper.tex
update to math
added a couple of comments

@@
  \dominique{it is not explained what cofes are?}
+ \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
 \begin{theorem}\label{thm:world-existence}
-  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and
+  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
   $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
@@
 
-The \emph{public future} region relation satisfies \dominique{is defined by?} the following properties:
+The \emph{public future} region relation satisfies
+ \dominique{is defined by?} 
+ \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations.
+It can be shown that these will have the properties mentioned below}
+the following properties:
 \begin{mathpar}
@@
     &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
-    &\qquad\begin{multlined}
-      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
-      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow 
-      \dom(\ms) = \dom(\ms') \text{\footnotesize{ and }} \\
-      \forall \addr \in
-      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
-    \end{multlined}
+    &\qquad
+      \begin{aligned}
+        &\forall s', \hat{W}, n, \ms, \ms' \ldotp \\
+        &\;\;\npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow\\
+        &\;\;\;\;\dom(\ms) = \dom(\ms') \text{\footnotesize{ and }} \\
+        &\;\;\;\;\forall \addr \in \dom(\ms)\ldotp\\
+        &\;\;\;\;\;\;\npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
+      \end{aligned}
   \end{align*}

### 2017-02-14 csf-2017/paper.tex
Two minor modifications.
Stack management added

@@
 contain arguments the adversary is intended to have access to. We leave a copy
-of the stack pointer in the appropriate register, but only after we have used
+of the stack pointer in $r_\stk$, but only after we have used
 the $\subseg{}{}{}$ to shrink its authority to just the part of the stack that
@@
 
-\textbf{Reusable macro instructions} With the high-level approach in mind, we
-define the following macro instructions: \texttt{fetch $r$ $\var{name}$} fetches
-the capability related to $\var{name}$ from the linking table and stores it in
-register $r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
-macros that uses the stack assumes a stack pointer in register $r_\stk$.
-\texttt{prepstk $r$} is used when a callback is invoked by the adversary and
-prepares the received stack pointer by checking that it has permission $\rwlx$
-and making it conform to our stack discipline. \texttt{scall
-  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps to the
-capability in register $r$ in the manner described above. That is, it pushes local state (the
-contents of registers $\overline{r_{\var{priv}}}$), stack pointer and return
-pointer to the stack, sets up the necessary on stack return code and creates a
-return pointer, restricts stack pointer, clears the unused part of the stack,
-clears the necessary registers and jumps to $r$. Upon return, the private state
-is restored. \texttt{mclear $r$} clears all the memory the capability in
-register $r$ has authority over. \texttt{rclear $\var{regSet}$} clears all the
-registers in $\var{regSet}$ \texttt{reqglob $r$} checks whether the word in
-register $r$ is a global capability. \texttt{crtcls $\overline{(x_i,r_i)}$ $r$}
-allocates a closure pair where the capability in register $r$ is used to point
-to the program and a new environment is allocated (using malloc) where the contents of
-$\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch
-happens when an instruction refers to $x_i$.
+\textbf{Stack management discipline}
+The precautions the stack discipline must take have now been introduced and motivated. Summarized, the stack discipline is:
+\begin{description}
+\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
+\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used.
+\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
+\item[Before invoking an adversary call] make sure it is a global capability.
+\item[When having a call back invoked by the adversary] make sure the alleged stack pointer has permission $\rwlx$.
+\end{description}
+
+\textbf{Reusable macro instructions} 
+We define a number of reusable macros. Some of them capture the stack
+management discipline and others make the examples in
+Section~\ref{sec:examples} readable. The macros are:
+\texttt{fetch $r$ $\var{name}$} fetches the capability related to
+$\var{name}$ from the linking table and stores it in register
+$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
+macros that uses the stack assumes a stack pointer in register
+$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
+the adversary and prepares the received stack pointer by checking that
+it has permission $\rwlx$ and making it conform to our stack
+discipline. \texttt{scall
+  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
+to the capability in register $r$ in the manner described above. That
+is, it pushes local state (the contents of registers
+$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
+stack, sets up the necessary on stack return code and creates a return
+pointer, restricts the stack pointer, clears the unused part of the
+stack, clears the necessary registers and jumps to $r$. Upon return,
+the private state is restored. \texttt{mclear $r$} clears all the
+memory the capability in register $r$ has authority
+over. \texttt{rclear $\var{regSet}$} clears all the registers in
+$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
+register $r$ is a global capability. \texttt{crtcls
+  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
+capability in register $r$ is used to point to the program and a new
+environment is allocated (using malloc) where the contents of
+$\overline{r_i}$ is stored. In the program referred to by $r$, an
+implicit fetch happens when an instruction refers to $x_i$.
 

### 2017-02-14 csf-2017/paper.tex
A shorter version of the stack management discipline paragraph

@@
                          ((\perm,\gl),\start,\addrend,\addr)\mid\\
-                         \start\in\Addrs,\addrend \in
+                         \start,\addr\in\Addrs,\addrend \in
                          \Addrs\cup \{\infty\}
@@
 The precautions the stack discipline must take have now been introduced and motivated. Summarized, the stack discipline is:
-\begin{description}
-\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
-\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used.
-\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
-\item[Before invoking an adversary call] make sure it is a global capability.
-\item[When having a call back invoked by the adversary] make sure the alleged stack pointer has permission $\rwlx$.
-\end{description}
+\emph{At program startup} a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
+\emph{Before returning to the adversary} clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used.
+\emph{Before invoking the adversary} push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
+\emph{Before invoking an adversary call} make sure it is a global capability.
+\emph{When having a call back invoked by the adversary} make sure the alleged stack pointer has permission $\rwlx$.
 
@@
 The interested reader can find all the technical details in the
-techincal appendix~\cite{XXX}.
+technical appendix~\cite{XXX}.
 

### 2017-02-14 csf-2017/paper.tex
Changed it back to the list version (for readability) we can change it back if we need the extra line

--- 2017-02-14 csf-2017/paper.tex
@@
 The precautions the stack discipline must take have now been introduced and motivated. Summarized, the stack discipline is:
-\begin{description}
+\begin{description}[font=\normalfont\itshape]
 \item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
@@
 \item[Before invoking an adversary call] make sure it is a global capability.
-\item[When having a call back invoked by the adversary] make sure the alleged stack pointer has permission $\rwlx$.
+\item[When having called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
 \end{description}
-
 \textbf{Reusable macro instructions} 

### 2017-02-14 csf-2017/paper.tex
Some clarifications for stack discipline flesh out.

@@
 \textbf{Stack management discipline}
-The precautions the stack discipline must take have now been introduced and motivated. Summarized, the stack discipline is:
+The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
 \begin{description}[font=\normalfont\itshape]
 \item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
-\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used.
+\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (if any part of the stack has been passed to an adversary, then that part must be erased).
 \item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
-\item[Before invoking an adversary call] make sure it is a global capability.
-\item[When having called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
+\item[Before invoking an adversary callback] make sure it is a global capability.
+\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
 \end{description}

### 2017-02-14 csf-2017/paper.tex
removed quantification from FTLR
Dealt with small comments in LR section

@@
 
-\dominique{``end up being defined recursively'' does not sound very insightful}
-The worlds end up being defined recursively: Worlds describe memory, which can
+%\dominique{``end up being defined recursively'' does not sound very insightful}
+The worlds are defined recursively: Worlds describe memory, which can
 contain capabilities, whose safety depends on worlds. We use the method of
@@
 % recursive domain equation.
- \dominique{it is not explained what cofes are?}
- \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
+% \dominique{it is not explained what cofes are?}
+% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
 \begin{theorem}\label{thm:world-existence}
@@
 The \emph{public future} region relation satisfies
- \dominique{is defined by?} 
- \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations.
+% \dominique{is defined by?} 
+% \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations.
 It can be shown that these will have the properties mentioned below}
@@
   \label{thm:ftlr}
-  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
+  If one of the following holds:
   \begin{align*}

### 2017-02-14 csf-2017/paper.tex
Some work on the examples section and discussion.

--- 2017-02-14 csf-2017/paper.tex
@@
 contain arguments the adversary is intended to have access to. We leave a copy
-of the stack pointer in $r_\stk$, but only after we have used
+of the stack pointer in the appropriate register, but only after we have used
 the $\subseg{}{}{}$ to shrink its authority to just the part of the stack that
@@
 
-\textbf{Stack management discipline}
-The precautions the stack discipline must take have now been introduced and motivated. Summarized, the stack discipline is:
-\begin{description}[font=\normalfont\itshape]
-\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
-\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used.
-\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
-\item[Before invoking an adversary call] make sure it is a global capability.
-\item[When having called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
-\end{description}
-\textbf{Reusable macro instructions} 
-We define a number of reusable macros. Some of them capture the stack
-management discipline and others make the examples in
-Section~\ref{sec:examples} readable. The macros are:
-\texttt{fetch $r$ $\var{name}$} fetches the capability related to
-$\var{name}$ from the linking table and stores it in register
-$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
-macros that uses the stack assumes a stack pointer in register
-$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
-the adversary and prepares the received stack pointer by checking that
-it has permission $\rwlx$ and making it conform to our stack
-discipline. \texttt{scall
-  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
-to the capability in register $r$ in the manner described above. That
-is, it pushes local state (the contents of registers
-$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
-stack, sets up the necessary on stack return code and creates a return
-pointer, restricts the stack pointer, clears the unused part of the
-stack, clears the necessary registers and jumps to $r$. Upon return,
-the private state is restored. \texttt{mclear $r$} clears all the
-memory the capability in register $r$ has authority
-over. \texttt{rclear $\var{regSet}$} clears all the registers in
-$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
-register $r$ is a global capability. \texttt{crtcls
-  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
-capability in register $r$ is used to point to the program and a new
-environment is allocated (using malloc) where the contents of
-$\overline{r_i}$ is stored. In the program referred to by $r$, an
-implicit fetch happens when an instruction refers to $x_i$.
+\textbf{Reusable macro instructions} With the high-level approach in mind, we
+define the following macro instructions: \texttt{fetch $r$ $\var{name}$} fetches
+the capability related to $\var{name}$ from the linking table and stores it in
+register $r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
+macros that uses the stack assumes a stack pointer in register $r_\stk$.
+\texttt{prepstk $r$} is used when a callback is invoked by the adversary and
+prepares the received stack pointer by checking that it has permission $\rwlx$
+and making it conform to our stack discipline. \texttt{scall
+  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps to the
+capability in register $r$ in the manner described above. That is, it pushes local state (the
+contents of registers $\overline{r_{\var{priv}}}$), stack pointer and return
+pointer to the stack, sets up the necessary on stack return code and creates a
+return pointer, restricts stack pointer, clears the unused part of the stack,
+clears the necessary registers and jumps to $r$. Upon return, the private state
+is restored. \texttt{mclear $r$} clears all the memory the capability in
+register $r$ has authority over. \texttt{rclear $\var{regSet}$} clears all the
+registers in $\var{regSet}$ \texttt{reqglob $r$} checks whether the word in
+register $r$ is a global capability. \texttt{crtcls $\overline{(x_i,r_i)}$ $r$}
+allocates a closure pair where the capability in register $r$ is used to point
+to the program and a new environment is allocated (using malloc) where the contents of
+$\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch
+happens when an instruction refers to $x_i$.
 
@@
 
-With the logical relation defined, we can now state a strong theorem
-about the safety properties of the capability machine:
-any capability that only grants reasonable authority is capability
-safe as a program counter.
+\subsection{Safety of the capability machine}
+With the logical relation defined, we can now state the fundamental theorem of
+our logical relation: a strong theorem that formalises the guarantees offered by
+the capability machine. Essentially, it says that a capability that only grants
+safe authority is capability safe as a program counter.
 \begin{theorem}[Fundamental Theorem]
@@
 \subsection{Encapsulation of local state}
-The capability machine provides encapsulation of local state. The
-programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
-\figurename~\ref{fig:prog-f2} demonstrate this.  The two programs
-\texttt{f1} and \texttt{f2} are very similar. They both store some
-local state, call an untrusted piece of code ($\adv$), and then tests
-whether the local state is unchanged. They differ in the way they do
-this. Program \texttt{f1} assumes it has access to a stack pointer
-and uses the stack to store its local state. It also uses the
-stack-based calling convention \texttt{scall} to call the untrusted code. On the
-other hand, \texttt{f2} allocates a new piece of memory for its local
-state and uses an activation-record based calling convention
-(described in the technical appendix) to run the adversarial code.
+The capability machine provides encapsulation of local state. The programs
+\texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
+\figurename~\ref{fig:prog-f2} demonstrate this. They are very similar: both
+store some local state, call an untrusted piece of code ($\adv$), and then test
+whether the local state is unchanged. They differ in the way they do this.
+Program \texttt{f1} assumes it has access to a stack pointer and uses the stack
+to store its local state. It also uses the stack-based calling convention
+\texttt{scall} to call the untrusted code. On the other hand, \texttt{f2} uses
+malloc to allocate memory for its local state and uses an activation-record
+based calling convention (described in the technical appendix) to run the
+adversarial code.
 
@@
 \end{lstlisting}
-  \caption{Simple program that uses a calling convention not relyant
+  \caption{Simple program that uses a calling convention not reliant
     on the stack}
@@
 % to make them simpler and a more complicated example is coming up.}
-The two examples also illustrate the versatility of the logical
-relation. The two examples use different calling convention, but the
-logical relation is not specific to any calling convention, so we can
-use it to reason about both programs.
+The two examples also illustrate the versatility of the logical relation.
+Because it is not specific to any calling convention, we can use it to reason
+about both programs, even though they use different calling conventions.
 
@@
 \]
-  where each of the capabilities have an appropriate range of authority and pointer furthermore
+  where each of the capabilities have an appropriate range of authority and pointer. Furthermore
   \begin{itemize}
@@
 
-To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the
-start configuration is capability safe (in the $\observations$
-relation) for a world with a permanent region that requires the
-assertion flag to be 0. This is done using an anti-reduction lemma
-which states that for a configuration to be capability safe, it is
-sufficient to show that it is in it at a later point of the
-execution. To reason about \texttt{scall}, it suffices to show that
-(1) the configuration just after we jumped is capability safe and (2)
-that the configuration just after \texttt{scall} is done cleaning up
-is capability safe.  We use the Fundamental Theorem to reason about
-the unknown code, but notice that the Fundamental Theorem says nothing
-about enter capabilities and that the adversary capability is an enter
-capability. Luckily the enter capability becomes a read/execute
-capability when we jump and the Fundamental Theorem can be used to
-show that this capability is capability safe as a program counter.
+To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
+configuration is capability safe (in the $\observations$ relation) for a world
+with a permanent region that requires the assertion flag to be 0. We prove this
+using an anti-reduction lemma by which it suffices to show that the
+configuration is safe after some reduction steps. We then use a general lemma
+for reasoning about \texttt{scall}, by which it suffices to show that (1) the
+configuration that \texttt{scall} will jump to is capability safe and (2) that
+the configuration just after \texttt{scall} is done cleaning up is capability
+safe. We use the Fundamental Theorem to reason about the unknown code, but
+notice that the Fundamental Theorem says nothing about enter capabilities and
+that the adversary capability is an enter capability. Luckily the enter
+capability becomes $\exec$ after the jump and then, the Fundamental Theorem
+applies.
 
@@
 
-\lau{in the following paragraph, we could make a reference to
-  \textbf{invoking the adversary} from
-  section~\ref{sec:stack-and-return-pointer} where the problem is
-  first introduced. In fact, this paragraph is a reiteration of the
-  problem in the aformentioned section.}
-In \texttt{f3} there are two calls to an adversary and in order for
-the assertion in the middle to succeed, they need to be
-well-bracketed. If the adversary is able to store the return
-pointer from the first call, so it is available for the second call, then it can be used
-in a context where \texttt{f3} has $2$ on top of its stack which will cause the
-assertion to fail. The return pointer is passed as a local
-capability, so the only means the adversary has to store it is the
-stack pointer. The stack pointer itself is a local capability,
-so the adversary can only store it on the stack. This means that after
-the adversary jumps to the return pointer, \texttt{f3} can
-revoke it (by erasing it which happens as part of the
-\texttt{scall}). In other words, the adversary cannot store the
-return pointer for the second call, and it is possible to
-prove that the assertion never fails. The correctness lemma for \texttt{f3} is similar to the two previous lemmas:
+In \texttt{f3} there are two calls to an adversary and in order for the
+assertion in the middle to succeed, they need to be well-bracketed. If the
+adversary were able to store the return pointer from the first call and invoke
+it in the second call, then \texttt{f3} would have $2$ on top of its stack and
+the assertion will fail. However, the security measures of our calling
+convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
+sort of attack: specifically, the return pointer is local, so it can only be
+stored on the stack, but the part of the stack that is accessible to the
+adversary is cleared before the second invocation. IThe following lemma shows
+that there are also no other attacks that will break well-bracketedness of this
+example, i.e. the assertion never fails. It is similar to the two previous
+lemmas:
 \begin{figure}[t]
@@
 
-The final example, in \figurename~\ref{fig:prog-g1}, is a faithful
-translation of a trciky example known from the
-litterature\cite{XXX}. It consists of two parts, \texttt{g1} and
-\texttt{f4}. \texttt{g1} is a closure generator that generates
-closures with one variable $x$ set to $0$ in its environment and
-\texttt{f4} as the program. The program \texttt{f4} expects two
-arguments: a stack and a callback. The stack is necessary to use
-\texttt{scall}, but because it comes from an untrusted source, we need
-to make sure that it has the correct permission and fits with our
-stack dicipline. The callback must be global for the program to work,
-we will get back to this point. The program now sets $x$ to $0$ and
-calls the callback. When it returns from the callback, it sets $x$ to
-$1$ and calls the callback a second time. When it returns from the
-callback, it asserts $x$ is $1$ and returns.
+The final example, in \figurename~\ref{fig:prog-g1}, is a faithful translation
+of a tricky example known from the
+literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
+parts, \texttt{g1} and \texttt{f4}. \texttt{g1} is a closure generator that
+generates closures with one variable $x$ set to $0$ in its environment and
+\texttt{f4} as the program. \texttt{f4} expects one argument, a callback. The
+program sets $x$ to $0$ and calls the callback. When it returns, $\texttt{f4}$
+sets $x$ to $1$ and calls the callback a second time. When it returns from the
+callback, it asserts that $x$ is $1$ and returns. This example is more
+complicated than the previous ones because it involves a closure invoked by the
+adversary and an adversary callback invoked by us. As explained in
+Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
+to properly check the stack pointer received by our closure from the adversary
+(it must have the correct permission and fit with our stack discipline) and (2)
+require that the adversary callback be global.
 
@@
 
-To illustrate how subtle this program is, consider one way an
-adversary may try to make the assertion fail. In the second callback
-an adversary can get to the first callback by calling the closure one
-more time. If the adversary can save the return pointer in
-the second callback and use it in the first, then the assertion will
-fail because $x$ is $0$ before the first callback. The return
-pointer cannot be transfered between the callbacks, so this does not
-work.
+To illustrate how subtle this program is, consider how an adversary could try to
+make the assertion fail. In the second callback an adversary can get to the
+first callback by invoking the closure one more time. If there were any way for
+the adversary to transfer the return pointer from the point where it reinvokes
+the closure to where the closure reinvokes the callback, then the assertion can
+be made to fail. Similarly, if there were any way for the adversary to store a
+stack pointer or trick the trusted code into preserving it across an invocation,
+the assertion can likely be made to fail too. However, our stack and return
+pointer discipline guarantees that there is no way any of this can happen, as we
+prove in the following lemma.
 
@@
 
-As mentioned, the callback must be global. This is to make sure the
-callback is not a piece of code on the stack that restores the
-adversaries original stack pointer breaking the encapsulation of
-our local stack frame. In the proof of Lemma~\ref{lem:correctness-g1},
-this also shows up. When we reason about \texttt{f4}, we revoke all
-the local state of the adversary which means that we consider a
-private future world. We use the $\execCond{}$ to reason about the
-jump to the callback and because the callback is global, it works for
-private future worlds. Had the callback been local, then the
-$\execCond{}$ would only work for public future worlds, and we would
-be stuck in the proof.
-
-In Lemma~\ref{lem:correctness-g1} the order of control has been
-inverted compared to the previous lemmas. In this lemma, the adversary
-assumes control first with a capability for the closure creator
-\texttt{g1}. This means we need to check all arguments are safe
-to use and that we clean up before returning in the end. The inversion
-of control poses an interesting challenge when it comes to reasoning
-about the adversaries local state during the execution of
-\texttt{f4}. During the execution of \texttt{f4} and the callbacks, the
-adversary should not be able to rely on its local state from before
-the call of \texttt{f4}. This is easily done by revoking all the
-temporary regions of the world given at the start of \texttt{f4}. When
-\texttt{f4} returns, then the adversary is allowed to rely on its
-local state again which means that we need to guarantee that the local
-state is unchanged. To illustrate how this is done in the proof of
-Lemma~\ref{lem:correctness-g1}, consider
-\figurename~\ref{fig:worlds-in-pf-g1}. The return pointer given to
-\texttt{f4} might be local and it is capability safe in $W_1$, so it
-can only be used in public future worlds. Let us see how $W_6$ is
-constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed by revoking all
-temporary regions and adding a $\iota^\pwl$ region for the stack we
-pass away and static temporary regions (a region that accepts exactly
-one memory) for our local stack as well as the adversaries local stack.
-The given worlds $W_3$ and $W_5$ are public future worlds. None of
-these worlds can have masked the temporary regions of $W_1$ with permanent
-regions. $W_6$ is constructed by revoking the temporary regions of
-$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$
-a public future world of $W_1$. It is therefore safe to use the return
-pointer to return to the adversary.
+As explained in Section~\ref{sec:stack-and-return-pointer}, the callback must be
+global, essentially to make sure that it is not allocated on the stack where it
+might contain the adversary's original stack pointer. If not, it could break the
+encapsulation of our local stack frame. In the proof of
+Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
+callback in a world that is only a private future world of the one where we
+received the callback, precisely because we have invalidated the local state of
+the adversary (particularly its previous stack and return capabilities). The
+callback is still valid in this private future world, but only because we know
+that it is global.
+
+In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
+compared to the previous lemmas. In this lemma, the adversary assumes control
+first with a capability for the closure creator \texttt{g1}. This means we need
+to check that all arguments are safe to use and that we clean up before
+returning in the end. The inversion of control poses an interesting challenge
+when it comes to reasoning about the adversary's local state during the
+execution of \texttt{f4}. During the execution of \texttt{f4} and the callbacks,
+the adversary should not be able to rely on its local state from before the call
+of \texttt{f4}. This is easily done by revoking all the temporary regions of the
+world given at the start of \texttt{f4}. However, when \texttt{f4} returns, the
+adversary is again allowed to rely on its old local state which means that we
+need to guarantee that the local state is unchanged. This is important because
+the return pointer that \texttt{f4} receives from the adversary may be local,
+and the adversary is allowed to allocate the activation record on the stack
+(just like we do) so that he can store and recover his old stack pointer after
+\texttt{f4} returns.
+
+In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accomodate this in the
+proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
+all the situations where control is passed to the adversary. The potentially
+local return pointer received by \texttt{f4} from the adversary is safe in
+$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
+constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
+by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
+we pass away and static temporary regions (a region that accepts exactly one
+memory) for our local stack as well as the adversaries local stack. The given
+worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
+to return to us. None of these worlds can have masked the temporary regions of
+$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
+regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
+$W_6$ a public future world of $W_1$. It is therefore safe to use the return
+pointer to return to the adversary, since we have restored validity of any local
+state he/she might have stashed away.
 
@@
 
-\begin{itemize}
-\item A general well-bracketed control flow result?
-  \begin{itemize}
-  \item what would that result say?
-  \item possible idea: fully abstract compilation from an assembly language with
-  a trusted stack to one without
-  \item the LR and some of the lemmas already imply well-bracketed control flow, as seen in examples
-  \end{itemize}
-\item Relation to local parameters in Scala, Algol/Pascal second-class function
-  parameters? 
-\item Stack clearing realistic?
-\item Non-modularity of heap allocation requirement for adversary callbacks
-\end{itemize}
-
-It is worth mentioning that local capabilities are quite similar to features
-that have been proposed for high-level languages. Specifically,
-\citet{osvald_gentrification_2016}'s second-class or local values in Scala are
-remarkably similar and we expect they will support similar reasoning patterns.
+\textbf{Formulating control flow integrity} The attentive reader may notice
+that, while we claim that our stack and return pointer discipline enforces
+control flow integrity, we do not prove a general theorem to show that this is
+the case. The reason for this is that it is not clear what such a theorem should
+look like. Formulations in terms of a control flow graph, like the one by
+\cite{abadi_control-flow_2005}, do not take into account temporal properties,
+like the well-bracketedness that Example~\texttt{g1} relies on.\footnote{It is a
+  bit unfortunate that the general term ``control flow integrity'' is used for a
+  property that only captures partial correctness of control flow.} In fact, our
+examples show that our logical relation imply a form of control flow integrity
+that is stronger than such formulations, although this is not made very
+explicit. As future work, we are considering to look at a more explicit and
+useful way to formalise well-bracketed control flow. The idea would be to define
+a variant of our capability machine with built-in call and return instructions
+and well-bracketed control flow built-in to the operational semantics, and then
+prove that compiling such code to our machine using our stack and return pointer
+discipline is fully abstract~\cite{abadi_protection_1998}, which roughly means
+that the lower-level machine does not allow interactions with the compiled code
+that do not correspond to an interaction with the original code in the
+high-level machine.
+
+\textbf{Performance and the requirement for stack clearing} While performance is
+not a goal of this
+
+\textbf{Modularity} Our return pointers and callbacks satisfy the checks we
+apply to the ones we get from the adversary (return pointers are allowed to be
+local, callbacks must be global), i.e. we do not claim special status and what
+we propose is modular.
+
+\textbf{Local capabilities in high-level languages} It is worth mentioning that
+local capabilities are quite similar to features that have been proposed for
+high-level languages. Specifically, \citet{osvald_gentrification_2016}'s
+second-class or local values in Scala are remarkably similar: they are a kind of
+values that can be provided to other code for immediate use without allowing it
+to store the value in a closure or reference for later use. We believe reasoning
+about such values will require very similar reasoning techniques than the ones
+we have used for local capabilities.
 

### 2017-02-14 csf-2017/paper.tex

@@
 
-\textbf{Reusable macro instructions} With the high-level approach in mind, we
-define the following macro instructions: \texttt{fetch $r$ $\var{name}$} fetches
-the capability related to $\var{name}$ from the linking table and stores it in
-register $r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
-macros that uses the stack assumes a stack pointer in register $r_\stk$.
-\texttt{prepstk $r$} is used when a callback is invoked by the adversary and
-prepares the received stack pointer by checking that it has permission $\rwlx$
-and making it conform to our stack discipline. \texttt{scall
-  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps to the
-capability in register $r$ in the manner described above. That is, it pushes local state (the
-contents of registers $\overline{r_{\var{priv}}}$), stack pointer and return
-pointer to the stack, sets up the necessary on stack return code and creates a
-return pointer, restricts stack pointer, clears the unused part of the stack,
-clears the necessary registers and jumps to $r$. Upon return, the private state
-is restored. \texttt{mclear $r$} clears all the memory the capability in
-register $r$ has authority over. \texttt{rclear $\var{regSet}$} clears all the
-registers in $\var{regSet}$ \texttt{reqglob $r$} checks whether the word in
-register $r$ is a global capability. \texttt{crtcls $\overline{(x_i,r_i)}$ $r$}
-allocates a closure pair where the capability in register $r$ is used to point
-to the program and a new environment is allocated (using malloc) where the contents of
-$\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch
-happens when an instruction refers to $x_i$.
+\textbf{Stack management discipline}
+The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
+\begin{description}[font=\normalfont\itshape]
+\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
+\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (if any part of the stack has been passed to an adversary, then that part must be erased).
+\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
+\item[Before invoking an adversary callback] make sure it is a global capability.
+\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
+\end{description}
+\textbf{Reusable macro instructions} 
+We define a number of reusable macros. Some of them capture the stack
+management discipline and others make the examples in
+Section~\ref{sec:examples} readable. The macros are:
+\texttt{fetch $r$ $\var{name}$} fetches the capability related to
+$\var{name}$ from the linking table and stores it in register
+$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
+macros that uses the stack assumes a stack pointer in register
+$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
+the adversary and prepares the received stack pointer by checking that
+it has permission $\rwlx$ and making it conform to our stack
+discipline. \texttt{scall
+  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
+to the capability in register $r$ in the manner described above. That
+is, it pushes local state (the contents of registers
+$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
+stack, sets up the necessary on stack return code and creates a return
+pointer, restricts the stack pointer, clears the unused part of the
+stack, clears the necessary registers and jumps to $r$. Upon return,
+the private state is restored. \texttt{mclear $r$} clears all the
+memory the capability in register $r$ has authority
+over. \texttt{rclear $\var{regSet}$} clears all the registers in
+$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
+register $r$ is a global capability. \texttt{crtcls
+  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
+capability in register $r$ is used to point to the program and a new
+environment is allocated (using malloc) where the contents of
+$\overline{r_i}$ is stored. In the program referred to by $r$, an
+implicit fetch happens when an instruction refers to $x_i$.
 

### 2017-02-14 csf-2017/paper.tex
stack management

--- 2017-02-14 csf-2017/paper.tex
@@
 \item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
-\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (if any part of the stack has been passed to an adversary, then that part must be erased).
-\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
+\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (any part of the stack that has been passed to an adversary must also be erased).
+\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in a register, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
 \item[Before invoking an adversary callback] make sure it is a global capability.

### 2017-02-14 csf-2017/paper.tex
pass over LR section

@@
 using an anti-reduction lemma by which it suffices to show that the
-configuration is safe after some reduction steps. We then use a general lemma
+configuration is capability safe after some reduction steps. We then use a general lemma
 for reasoning about \texttt{scall}, by which it suffices to show that (1) the
@@
 assertion in the middle to succeed, they need to be well-bracketed. If the
-adversary were able to store the return pointer from the first call and invoke
+adversary was able to store the return pointer from the first call and invoke
 it in the second call, then \texttt{f3} would have $2$ on top of its stack and
@@
 stored on the stack, but the part of the stack that is accessible to the
-adversary is cleared before the second invocation. IThe following lemma shows
+adversary is cleared before the second invocation. The following lemma shows
 that there are also no other attacks that will break well-bracketedness of this
@@
 (it must have the correct permission and fit with our stack discipline) and (2)
-require that the adversary callback be global.
+require that the adversary callback is global.
 
@@
 and the adversary is allowed to allocate the activation record on the stack
-(just like we do) so that he can store and recover his old stack pointer after
+(just like we do) so that they can store and recover his old stack pointer after
 \texttt{f4} returns.
@@
 we pass away and static temporary regions (a region that accepts exactly one
-memory) for our local stack as well as the adversaries local stack. The given
+memory) for our local stack as well as the adversary's local stack. The given
 worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
@@
 pointer to return to the adversary, since we have restored validity of any local
-state he/she might have stashed away.
+state they might have stashed away.
 

### 2017-02-14 csf-2017/paper.tex
Removed formal definitions of standard islands and added a short in text description (all in the name of more space)

@@
 \end{itemize}
-\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution.
+%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution.
 Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}
@@
 
-\begin{figure}[htbp]
-  \centering
-  \begin{align*}
-  &\iota_{\start,\addrend}^\pwl : \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
-  \\
-  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
-    \begin{aligned}
-      & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
-      &\forall \addr \in [\start,\addrend] \ldotp\\
-      & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-    \end{aligned}
-        \right\}\\
-  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
-  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
-  \\
-  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
-    \begin{aligned}
-      &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
-      &\forall \addr \in [\start,\addrend] \ldotp \\
-      & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
-      & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-     \end{aligned}
-     \right\}
-\end{align*}
-\caption{Standard regions}
-\label{fig:standard-regions}
-\end{figure}
+% \begin{figure}[htbp]
+%   \centering
+%   \begin{align*}
+%   &\iota_{\start,\addrend}^\pwl : \Regions\\
+%   &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
+%   \\
+%   &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
+%   &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
+%     \begin{aligned}
+%       & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
+%       &\forall \addr \in [\start,\addrend] \ldotp\\
+%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
+%     \end{aligned}
+%         \right\}\\
+%   &\iota_{\start,\addrend}^{\nwl} : \Regions \\
+%   &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
+%   \\
+%   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
+%   &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
+%     \begin{aligned}
+%       &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
+%       &\forall \addr \in [\start,\addrend] \ldotp \\
+%       & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
+%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
+%      \end{aligned}
+%      \right\}
+% \end{align*}
+% \caption{Standard regions}
+% \label{fig:standard-regions}
+% \end{figure}
 
@@
 intuition why they define capability safety. The definition of the logical
-relation is found in \figurename~\ref{fig:logrel},
-\figurename~\ref{fig:standard-regions}, and \figurename~\ref{fig:perm-conds}.
+relation is found in \figurename~\ref{fig:logrel}, and \figurename~\ref{fig:perm-conds}.
 Some definitions are omitted and explained when needed. The precise definitions
@@
 depend on permanent regions. Therefore we use the function
-$\var{localityReg}(g,W)$, which projects out all active regions when $g$ is
-local, but only the permanent regions when $g$ is global. The read condition
-uses $\iota^\pwl_{\start,\addrend}$, a standard regions defined in
-Figure~\ref{fig:standard-regions}, which requires all the words in the range
-$[\start,\addrend]$ are safe.
-
-For a capability with write permission, $\writeCond{}$ must be satisfied for the
-capability's range of authority. Such a capability may be used to write any
-word, so the region governing the relevant part of memory must (at least) allow
-any capability safe word to be stored. However, this is not entirely true
-because of the distinction between write and permit-write-local capabilities.
-Only permit-write-local capabilities need to allow arbitrary safe values to be
-stored, while for write capabilities, it is sufficient to allow storing values
-that are both non-local and safe. The write condition uses $\var{localityReg}$
-to pick an appropriate region for its locality in the same way the read
-condition does. The condition for the region to be address-stratified is there
-for technical reasons.
+$\var{localityReg}(\gl,W)$, which projects out all active regions when
+$\gl$ (the locality) is local, but only the permanent regions when $\gl$ is global.
+The read condition is defined in terms of a standard region
+$\iota^\pwl_{\start,\addrend}$ which requires all the words in the range
+$[\start,\addrend]$ to be safe.
+
+For a capability with write permission, $\writeCond{}$ must be
+satisfied for the capability's range of authority. Such a capability
+may be used to write any word, so the region governing the relevant
+part of memory must (at least) allow any capability safe word to be
+stored. However, this is not entirely true because of the distinction
+between write and permit-write-local capabilities.  Only
+permit-write-local capabilities need to allow arbitrary safe values to
+be stored, while for write capabilities, it is sufficient to allow
+storing values that are both non-local and safe. The write condition
+uses $\var{localityReg}$ to pick an appropriate region for its
+locality in the same way the read condition does. The condition for
+the region to be address-stratified is there for technical reasons. In
+the value relation, the write condition is used with $\iota^\pwl$ for
+write-local capabilities. For write capabilities, the write condition
+is used with another standard region $\iota^\nwl_{\start,\addrend}$
+which requires all the words in the range $[\start,\addrend]$ to be
+non-local and safe.
 

### 2017-02-14 csf-2017/paper.tex
normalised 'i.e'

@@
 into the appropriate register, but only if the capability has appropriate
-authority (i.e. read permission and an appropriate range). The $\restricttwo{}{}$
+authority (i.e.\ read permission and an appropriate range). The $\restricttwo{}{}$
 instruction will update a capability's permissions and global/local status in
@@
 CHERI~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
-they can only be kept in registers, i.e. they cannot be stored to memory. This
+they can only be kept in registers, i.e.\ they cannot be stored to memory. This
 means that local capabilities can be \emph{temporarily} given to an adversary,
@@
 to the rule above: local capabilities can be stored to a part of memory for
-which we have a capability with write-local authority (i.e. permission $\rwl$ or
+which we have a capability with write-local authority (i.e.\ permission $\rwl$ or
 $\rwlx$). This exception is intended to accomodate regions of memory that behave
@@
 runtime to enforce this by not handing out such capabilities for other parts of
-memory in any situation (i.e. not through malloc, not during loading or linking
+memory in any situation (i.e.\ not through malloc, not during loading or linking
 etc.). Additionally, in our examples, we typically also assume some static parts
@@
 for allocated regions as requested, but it will never hand out capabilities with
-write-local capability (i.e. no $\rwlx$ or $\rwl$ permissions). For simplicity,
+write-local capability (i.e.\ no $\rwlx$ or $\rwl$ permissions). For simplicity,
 we assume that memory allocated through malloc cannot be freed.
@@
 before we do (erasing any data we do not want to pass as (part of) the return
-value) and similarly, erase any part of the stack that we have used (i.e.
+value) and similarly, erase any part of the stack that we have used (i.e.\
 overwrite it with zeros). Note that we can use the stack as normal: we may push
@@
 A final required safety measure is that before invocation of the adversary, we
-need to clear (i.e. overwrite with zeros) the part of the stack that we are not
+need to clear (i.e.\ overwrite with zeros) the part of the stack that we are not
 using. Otherwise, that part of the stack may still contain data and capabilities
@@
 adversary. In other words, the return pointer provided to the adversary must be
-a capability that they can jump to but not read from, i.e. an enter
+a capability that they can jump to but not read from, i.e.\ an enter
 capability. To make this work, we construct the activation record as depicted in
@@
 local, the activation record can only be constructed in a part of memory where
-we have write-local access, i.e. on the stack. Note that this means we will be
-placing and executing instructions on the stack, i.e. it will not just contain
+we have write-local access, i.e.\ on the stack. Note that this means we will be
+placing and executing instructions on the stack, i.e.\ it will not just contain
 code pointers and data. This means that our discipline should be combined with
-protection against stack smashing attacks (i.e. buffer overflows on the stack
+protection against stack smashing attacks (i.e.\ buffer overflows on the stack
 overwriting activation records' contents), but conveniently, the fine-grained
@@
 our local stack frame during a callback by passing us a local $\rwx$
-capability to a global data structure (i.e., there exists a global
+capability to a global data structure (i.e.\ there exists a global
 capability for it) as the stack pointer and a global
@@
 The \emph{register-file relation} $\stdrr$ specifies that register-files are
-safe when they contain safe words (i.e., words in $\stdvr$) in all registers but
+safe when they contain safe words (i.e.\ words in $\stdvr$) in all registers but
 $\pcreg$. The \emph{expression relation} $\stder$ defines when a word is safe
 to use as a program counter. That is the case when the word can be plugged into
-a capability safe register file (i.e., a register file in $\stdrr$) and paired
+a capability safe register file (i.e.\ a register file in $\stdrr$) and paired
 with a memory satisfying the world to become a capability safe configuration.
 
-The \emph{value relation} $\stdvr$ defines when words are safe, i.e. when they
+The \emph{value relation} $\stdvr$ defines when words are safe, i.e.\ when they
 cannot be used to break the protocols defined by the world. Non-capability data
@@
 that there are also no other attacks that will break well-bracketedness of this
-example, i.e. the assertion never fails. It is similar to the two previous
+example, i.e.\ the assertion never fails. It is similar to the two previous
 lemmas:
@@
 apply to the ones we get from the adversary (return pointers are allowed to be
-local, callbacks must be global), i.e. we do not claim special status and what
+local, callbacks must be global), i.e.\ we do not claim special status and what
 we propose is modular.

### 2017-02-14 csf-2017/paper.tex
Added another version of figure 3 (in comment). Saves a bit of space, but is not as clear and lacks the 'failed' case of table
typeset programs so they have same font size as math
updated in text code to match the new style

@@
 \lstset{
-  basicstyle=\ttfamily,
+  basicstyle=\footnotesize\ttfamily,
   columns=fullflexible,
@@
 
+\begin{comment}
+  \begin{figure}
+    \centering
+    \begin{mathpar}
+      \Phi \rightarrow
+      \begin{cases}
+        \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
+        \begin{array}[t]{l}
+          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\
+          \;\;\text{\footnotesize{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}}\\
+          \;\;\text{\footnotesize{and $\perm \in \{ \exec,\rwx, \rwlx \}$} }
+        \end{array}\\
+        \failed                & \text{\footnotesize{otherwise}}
+      \end{cases}
+      \and \updatePcPerm{w} =
+      \begin{cases}
+        ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
+        w & \text{\footnotesize{otherwise}}
+      \end{cases}\\
+      \and \stdUpdatePc{\Phi} =
+      \begin{cases}
+        \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
+        \begin{array}[t]{l}
+          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\ 
+          \;\;\text{\footnotesize{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}
+        \end{array}\\
+        \failed & \text{\footnotesize{otherwise}}
+      \end{cases}
+      \and \inferrule{ } { \sem{\fail}(\Phi) = \failed } \and
+      \inferrule{ } { \sem{\halt}(\Phi) = (\halted,\memheap) } \and
+      \inferrule{ } { \sem{\move{\refreg{r_1}}{r_2}}(\Phi) =
+        \stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)}) } \and
+      \inferrule{ \memreg(r_2) = \stdcap \\
+        \perm \in \{ \rwx, \rwlx, \exec, \readwrite,
+        \readwritel, \readonly \}\\
+        \start \leq \addr \leq \addrend } {
+        \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi) =
+        \updateReg{r_1}{\memheap(\addr)}} \and
+      \inferrule{ \memreg(r_2) = \stdcap[\permp]\\
+        \permp' = \decodePermPair{\memreg(r_2)}\\
+        \permp'\sqsubseteq \permp } {
+        \sem{\restricttwo{\refreg{r_1}}{r_2}}(\Phi) = \\
+        \stdUpdatePc{}(\updateReg{r_1}{(\permp',\start,\addrend,\addr)})
+      } \and \inferrule{ \memreg(r_2) = ((\_,\_),\_,\_,\addr) } {
+        \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) =
+        \stdUpdatePc{\updateReg{r_1}{\addr}}} \and \inferrule{
+        \var{newPc} = \updatePcPerm{\memreg(\lv)} } {
+        \sem{\jmp{\lv}}(\Phi) = \updateReg{\pcreg}{\var{newPc}} } \and
+      \inferrule{ \memreg(r_1) = \stdcap\\
+        \start \leq \addr \leq \addrend\\
+        \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}\\
+        \var{w} = \memreg(r_2) \\
+        \var{w} = ((\_,\local),\_,\_,\_)\Rightarrow\perm \in
+        \{\rwlx,\readwritel \} } {
+        \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) =
+        \stdUpdatePc{\updateHeap{\addr}{\var{w}}} } \and
+    \end{mathpar}
+    \caption{An excerpt from the operational semantics.}
+    \label{fig:op-sem}
+  \end{figure}
+\end{comment}
 \figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics containing a number of representative
@@
 Section~\ref{sec:examples} readable. The macros are:
-\texttt{fetch $r$ $\var{name}$} fetches the capability related to
+\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
 $\var{name}$ from the linking table and stores it in register
-$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
+$r$. \texttt{\footnotesize{push $r$}}/\texttt{\footnotesize{pop $r$}} works as one would expect. All
 macros that uses the stack assumes a stack pointer in register
-$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
+$r_\stk$.  \texttt{\footnotesize{prepstk $r$}} is used when a callback is invoked by
 the adversary and prepares the received stack pointer by checking that
 it has permission $\rwlx$ and making it conform to our stack
-discipline. \texttt{scall
-  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
+discipline. \texttt{\footnotesize{scall $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} Jumps
 to the capability in register $r$ in the manner described above. That
@@
 stack, clears the necessary registers and jumps to $r$. Upon return,
-the private state is restored. \texttt{mclear $r$} clears all the
+the private state is restored. \texttt{\footnotesize{mclear $r$}} clears all the
 memory the capability in register $r$ has authority
-over. \texttt{rclear $\var{regSet}$} clears all the registers in
-$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
-register $r$ is a global capability. \texttt{crtcls
-  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
+over. \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
+$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
+register $r$ is a global capability. \texttt{\footnotesize{crtcls
+  $\overline{(x_i,r_i)}$ $r$}} allocates a closure pair where the
 capability in register $r$ is used to point to the program and a new
@@
 The capability machine provides encapsulation of local state. The programs
-\texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
+\texttt{\footnotesize{f1}} in \figurename~\ref{fig:prog-f1} and \texttt{\footnotesize{f2}}
 \figurename~\ref{fig:prog-f2} demonstrate this. They are very similar: both
@@
 whether the local state is unchanged. They differ in the way they do this.
-Program \texttt{f1} assumes it has access to a stack pointer and uses the stack
+Program \texttt{\footnotesize{f1}} assumes it has access to a stack pointer and uses the stack
 to store its local state. It also uses the stack-based calling convention
-\texttt{scall} to call the untrusted code. On the other hand, \texttt{f2} uses
+\texttt{\footnotesize{scall}} to call the untrusted code. On the other hand, \texttt{\footnotesize{f2}} uses
 malloc to allocate memory for its local state and uses an activation-record
@@
 
-In order to formulate lemmas about \texttt{f1} and \texttt{f2}, we
+In order to formulate lemmas about \texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}}, we
 need a way to observe whether the assertion fails. To this end, we
@@
 fails, it sets the flag to $1$ and halts. A slightly simplified
-version of the correctness lemma for \texttt{f1} is:
+version of the correctness lemma for \texttt{\footnotesize{f1}} is:
 \begin{lemma}
@@
   \begin{itemize}
-  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
+  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
   \item $\ms_\flag(\flag) = 0$
@@
 configuration is capability safe after some reduction steps. We then use a general lemma
-for reasoning about \texttt{scall}, by which it suffices to show that (1) the
-configuration that \texttt{scall} will jump to is capability safe and (2) that
-the configuration just after \texttt{scall} is done cleaning up is capability
+for reasoning about \texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
+configuration that \texttt{\footnotesize{scall}} will jump to is capability safe and (2) that
+the configuration just after \texttt{\footnotesize{scall}} is done cleaning up is capability
 safe. We use the Fundamental Theorem to reason about the unknown code, but
@@
 
-We have a similar lemma for \texttt{f2}:
+We have a similar lemma for \texttt{\footnotesize{f2}}:
 \begin{lemma}
@@
 \subsection{Well-bracketed control-flow} 
-Using the stack-based calling convention of \texttt{scall}, we get
+Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
 well-bracketed control-flow. To illustrate this, we look at
-two example programs \texttt{f3} in \figurename~\ref{fig:prog-f3} and
-\texttt{g1} in \figurename~\ref{fig:prog-g1}.
-
-In \texttt{f3} there are two calls to an adversary and in order for the
+two example programs \texttt{\footnotesize{f3}} in \figurename~\ref{fig:prog-f3} and
+\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-g1}.
+
+In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order for the
 assertion in the middle to succeed, they need to be well-bracketed. If the
 adversary was able to store the return pointer from the first call and invoke
-it in the second call, then \texttt{f3} would have $2$ on top of its stack and
+it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on top of its stack and
 the assertion will fail. However, the security measures of our calling
@@
 literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
-parts, \texttt{g1} and \texttt{f4}. \texttt{g1} is a closure generator that
+parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
 generates closures with one variable $x$ set to $0$ in its environment and
-\texttt{f4} as the program. \texttt{f4} expects one argument, a callback. The
-program sets $x$ to $0$ and calls the callback. When it returns, $\texttt{f4}$
+\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
+program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
 sets $x$ to $1$ and calls the callback a second time. When it returns from the
@@
     \texttt{scall}s to function correctly. $\var{offset}$ is the
-    offset to $f4$.}
+    offset to \texttt{f4}.}
   \label{fig:prog-g1}
@@
 compared to the previous lemmas. In this lemma, the adversary assumes control
-first with a capability for the closure creator \texttt{g1}. This means we need
+first with a capability for the closure creator \texttt{\footnotesize{g1}}. This means we need
 to check that all arguments are safe to use and that we clean up before
@@
 when it comes to reasoning about the adversary's local state during the
-execution of \texttt{f4}. During the execution of \texttt{f4} and the callbacks,
+execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
 the adversary should not be able to rely on its local state from before the call
-of \texttt{f4}. This is easily done by revoking all the temporary regions of the
-world given at the start of \texttt{f4}. However, when \texttt{f4} returns, the
+of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
+world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
 adversary is again allowed to rely on its old local state which means that we
 need to guarantee that the local state is unchanged. This is important because
-the return pointer that \texttt{f4} receives from the adversary may be local,
+the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
 and the adversary is allowed to allocate the activation record on the stack
 (just like we do) so that they can store and recover his old stack pointer after
-\texttt{f4} returns.
+\texttt{\footnotesize{f4}} returns.
 
@@
 all the situations where control is passed to the adversary. The potentially
-local return pointer received by \texttt{f4} from the adversary is safe in
+local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
 $W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
@@
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
-like the well-bracketedness that Example~\texttt{g1} relies on.\footnote{It is a
+like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.\footnote{It is a
   bit unfortunate that the general term ``control flow integrity'' is used for a

### 2017-02-14 csf-2017/paper.tex
removed definition of address-stratisfied and referred to the technical appendix in the text
combined f1 and f2 figure into one

@@
   \end{align*}
-  \begin{align*}
-    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
-    &\qquad
-      \begin{aligned}
-        &\forall s', \hat{W}, n, \ms, \ms' \ldotp \\
-        &\;\;\npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow\\
-        &\;\;\;\;\dom(\ms) = \dom(\ms') \text{\footnotesize{ and }} \\
-        &\;\;\;\;\forall \addr \in \dom(\ms)\ldotp\\
-        &\;\;\;\;\;\;\npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
-      \end{aligned}
-  \end{align*}
-
 \caption{Permission-based conditions}
@@
 locality in the same way the read condition does. The condition for
-the region to be address-stratified is there for technical reasons. In
+the region to be address-stratified is there for technical reasons and for the full definition, we refer to the technical appendix. In
 the value relation, the write condition is used with $\iota^\pwl$ for
@@
 The capability machine provides encapsulation of local state. The programs
-\texttt{\footnotesize{f1}} in \figurename~\ref{fig:prog-f1} and \texttt{\footnotesize{f2}}
-\figurename~\ref{fig:prog-f2} demonstrate this. They are very similar: both
+\texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
+\figurename~\ref{fig:prog-f1-and-f2} demonstrate this. They are very similar: both
 store some local state, call an untrusted piece of code ($\adv$), and then test
@@
 \end{lstlisting}
-  \caption{Simple program that uses a stack-based calling convention.}
-  \label{fig:prog-f1}
-\end{figure}
-
-\begin{figure}[t]
-  \centering
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-  \caption{Simple program that uses a calling convention not reliant
-    on the stack}
-  \label{fig:prog-f2}
+  \caption{Two simple programs that rely on local-state encapsulation. \texttt{f1} uses our stack-based calling convention. \texttt{f2} uses a calling not reliant on the stack}
+  \label{fig:prog-f1-and-f2}
 \end{figure}

### 2017-02-14 csf-2017/paper.tex
udpated f1 and f2 paper to have the programs side by side
Add a discussion of stack clearing and fix a missing requirement in the stack discipline: stack needs to be cleared before returning...

--- 2017-02-14 csf-2017/paper.tex
@@
 \end{itemize}
-%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution.
+\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution.
 Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}
@@
 into the appropriate register, but only if the capability has appropriate
-authority (i.e.\ read permission and an appropriate range). The $\restricttwo{}{}$
+authority (i.e. read permission and an appropriate range). The $\restricttwo{}{}$
 instruction will update a capability's permissions and global/local status in
@@
 CHERI~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
-they can only be kept in registers, i.e.\ they cannot be stored to memory. This
+they can only be kept in registers, i.e. they cannot be stored to memory. This
 means that local capabilities can be \emph{temporarily} given to an adversary,
@@
 to the rule above: local capabilities can be stored to a part of memory for
-which we have a capability with write-local authority (i.e.\ permission $\rwl$ or
+which we have a capability with write-local authority (i.e. permission $\rwl$ or
 $\rwlx$). This exception is intended to accomodate regions of memory that behave
@@
 runtime to enforce this by not handing out such capabilities for other parts of
-memory in any situation (i.e.\ not through malloc, not during loading or linking
+memory in any situation (i.e. not through malloc, not during loading or linking
 etc.). Additionally, in our examples, we typically also assume some static parts
@@
 for allocated regions as requested, but it will never hand out capabilities with
-write-local capability (i.e.\ no $\rwlx$ or $\rwl$ permissions). For simplicity,
+write-local capability (i.e. no $\rwlx$ or $\rwl$ permissions). For simplicity,
 we assume that memory allocated through malloc cannot be freed.
@@
 before we do (erasing any data we do not want to pass as (part of) the return
-value) and similarly, erase any part of the stack that we have used (i.e.\
-overwrite it with zeros). Note that we can use the stack as normal: we may push
-values onto it and pop them back off (e.g. for register spilling etc.).
+value) and similarly. However, it is important to carefully consider what needs
+to happen with the stack. Since we may have used the stack for various purposes
+(register spilling, storing local state when invoking other functions etc.), we
+clearly need to clear that data before returning to the adversary, to prevent
+the adversary from gaining access to it.
+
+However, if we are returning from a function that has itself invoked adversary
+code, then even the unused part of the stack may still contain data and
+capabilities left there by them, including local capabilities since the stack is
+store-local. The adversary could take advantage of this by using the unused part
+of the stack to store either stack pointers or return pointers from such
+previous invocations and then use them when they're no longer supposed to. To
+prevent this, before returning, we need to clear (i.e. overwrite with zeros) the
+entire part of the stack that the adversary has had access to, not just the
+parts that we have used ourselves. Essentially, this is needed for the same
+reason that we clear the register file: because the adversary should not be able
+to keep hold of local capabilities when control is passed to the trusted code
+and back to the adversary. Clearing the entire unused part of the stack is the
+most problematic aspect of our discipline in terms of efficiency, since it may
+consist of a large part of memory. However, we believe there are ways to
+optimize it that we discuss in Section~\ref{sec:discussion}.
 
@@
 A final required safety measure is that before invocation of the adversary, we
-need to clear (i.e.\ overwrite with zeros) the part of the stack that we are not
-using. Otherwise, that part of the stack may still contain data and capabilities
-from previously executed code, including local capabilities since they are
-allowed on the stack. The adversary could take advantage of this by using it to
-store either stack pointers or return pointers from previous invocations and
-play the same tricks as above. Clearing the entire unused part of the stack is
-the most problematic aspect of our discipline in terms of efficiency, since it
-may consist of a large part of memory. We believe that capability machines like
-CHERI should consider to provide special support for this requirement, by
-implementing an optimized instruction for clearing
-a large block of memory. We have no experience in processor design, but we
-imagine this might be optimized as a vector instruction or by having the
-processor keep track of a range of memory for which loads should return zero
-irrespective of the actual contents of the memory.
+need to clear the part of the stack that we are not using. Otherwise, similarly
+to what we explained above for returning to the adversary, that part of the
+stack may still contain local capabilities from previously executed code and the
+adversary could exploit this in much the same way as above.
 
@@
 adversary. In other words, the return pointer provided to the adversary must be
-a capability that they can jump to but not read from, i.e.\ an enter
+a capability that they can jump to but not read from, i.e. an enter
 capability. To make this work, we construct the activation record as depicted in
@@
 local, the activation record can only be constructed in a part of memory where
-we have write-local access, i.e.\ on the stack. Note that this means we will be
-placing and executing instructions on the stack, i.e.\ it will not just contain
+we have write-local access, i.e. on the stack. Note that this means we will be
+placing and executing instructions on the stack, i.e. it will not just contain
 code pointers and data. This means that our discipline should be combined with
-protection against stack smashing attacks (i.e.\ buffer overflows on the stack
+protection against stack smashing attacks (i.e. buffer overflows on the stack
 overwriting activation records' contents), but conveniently, the fine-grained
@@
 our local stack frame during a callback by passing us a local $\rwx$
-capability to a global data structure (i.e.\ there exists a global
+capability to a global data structure (i.e., there exists a global
 capability for it) as the stack pointer and a global
@@
 \item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
-\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (any part of the stack that has been passed to an adversary must also be erased).
-\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in a register, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
+\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (if any part of the stack has been passed to an adversary, then that part must be erased).
+\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
 \item[Before invoking an adversary callback] make sure it is a global capability.
@@
 
-% \begin{figure}[htbp]
-%   \centering
-%   \begin{align*}
-%   &\iota_{\start,\addrend}^\pwl : \Regions\\
-%   &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
-%   \\
-%   &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-%   &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
-%     \begin{aligned}
-%       & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
-%       &\forall \addr \in [\start,\addrend] \ldotp\\
-%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-%     \end{aligned}
-%         \right\}\\
-%   &\iota_{\start,\addrend}^{\nwl} : \Regions \\
-%   &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
-%   \\
-%   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-%   &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
-%     \begin{aligned}
-%       &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
-%       &\forall \addr \in [\start,\addrend] \ldotp \\
-%       & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
-%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-%      \end{aligned}
-%      \right\}
-% \end{align*}
-% \caption{Standard regions}
-% \label{fig:standard-regions}
-% \end{figure}
+\begin{figure}[htbp]
+  \centering
+  \begin{align*}
+  &\iota_{\start,\addrend}^\pwl : \Regions\\
+  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
+  \\
+  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
+  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
+    \begin{aligned}
+      & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
+      &\forall \addr \in [\start,\addrend] \ldotp\\
+      & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
+    \end{aligned}
+        \right\}\\
+  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
+  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
+  \\
+  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
+  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
+    \begin{aligned}
+      &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
+      &\forall \addr \in [\start,\addrend] \ldotp \\
+      & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
+      & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
+     \end{aligned}
+     \right\}
+\end{align*}
+\caption{Standard regions}
+\label{fig:standard-regions}
+\end{figure}
 
@@
 intuition why they define capability safety. The definition of the logical
-relation is found in \figurename~\ref{fig:logrel}, and \figurename~\ref{fig:perm-conds}.
+relation is found in \figurename~\ref{fig:logrel},
+\figurename~\ref{fig:standard-regions}, and \figurename~\ref{fig:perm-conds}.
 Some definitions are omitted and explained when needed. The precise definitions
@@
 The \emph{register-file relation} $\stdrr$ specifies that register-files are
-safe when they contain safe words (i.e.\ words in $\stdvr$) in all registers but
+safe when they contain safe words (i.e., words in $\stdvr$) in all registers but
 $\pcreg$. The \emph{expression relation} $\stder$ defines when a word is safe
 to use as a program counter. That is the case when the word can be plugged into
-a capability safe register file (i.e.\ a register file in $\stdrr$) and paired
+a capability safe register file (i.e., a register file in $\stdrr$) and paired
 with a memory satisfying the world to become a capability safe configuration.
 
-The \emph{value relation} $\stdvr$ defines when words are safe, i.e.\ when they
+The \emph{value relation} $\stdvr$ defines when words are safe, i.e. when they
 cannot be used to break the protocols defined by the world. Non-capability data
@@
 depend on permanent regions. Therefore we use the function
-$\var{localityReg}(\gl,W)$, which projects out all active regions when
-$\gl$ (the locality) is local, but only the permanent regions when $\gl$ is global.
-The read condition is defined in terms of a standard region
-$\iota^\pwl_{\start,\addrend}$ which requires all the words in the range
-$[\start,\addrend]$ to be safe.
-
-For a capability with write permission, $\writeCond{}$ must be
-satisfied for the capability's range of authority. Such a capability
-may be used to write any word, so the region governing the relevant
-part of memory must (at least) allow any capability safe word to be
-stored. However, this is not entirely true because of the distinction
-between write and permit-write-local capabilities.  Only
-permit-write-local capabilities need to allow arbitrary safe values to
-be stored, while for write capabilities, it is sufficient to allow
-storing values that are both non-local and safe. The write condition
-uses $\var{localityReg}$ to pick an appropriate region for its
-locality in the same way the read condition does. The condition for
-the region to be address-stratified is there for technical reasons. In
-the value relation, the write condition is used with $\iota^\pwl$ for
-write-local capabilities. For write capabilities, the write condition
-is used with another standard region $\iota^\nwl_{\start,\addrend}$
-which requires all the words in the range $[\start,\addrend]$ to be
-non-local and safe.
+$\var{localityReg}(g,W)$, which projects out all active regions when $g$ is
+local, but only the permanent regions when $g$ is global. The read condition
+uses $\iota^\pwl_{\start,\addrend}$, a standard regions defined in
+Figure~\ref{fig:standard-regions}, which requires all the words in the range
+$[\start,\addrend]$ are safe.
+
+For a capability with write permission, $\writeCond{}$ must be satisfied for the
+capability's range of authority. Such a capability may be used to write any
+word, so the region governing the relevant part of memory must (at least) allow
+any capability safe word to be stored. However, this is not entirely true
+because of the distinction between write and permit-write-local capabilities.
+Only permit-write-local capabilities need to allow arbitrary safe values to be
+stored, while for write capabilities, it is sufficient to allow storing values
+that are both non-local and safe. The write condition uses $\var{localityReg}$
+to pick an appropriate region for its locality in the same way the read
+condition does. The condition for the region to be address-stratified is there
+for technical reasons.
 
@@
 using an anti-reduction lemma by which it suffices to show that the
-configuration is capability safe after some reduction steps. We then use a general lemma
+configuration is safe after some reduction steps. We then use a general lemma
 for reasoning about \texttt{scall}, by which it suffices to show that (1) the
@@
 assertion in the middle to succeed, they need to be well-bracketed. If the
-adversary was able to store the return pointer from the first call and invoke
+adversary were able to store the return pointer from the first call and invoke
 it in the second call, then \texttt{f3} would have $2$ on top of its stack and
@@
 stored on the stack, but the part of the stack that is accessible to the
-adversary is cleared before the second invocation. The following lemma shows
+adversary is cleared before the second invocation. IThe following lemma shows
 that there are also no other attacks that will break well-bracketedness of this
-example, i.e.\ the assertion never fails. It is similar to the two previous
+example, i.e. the assertion never fails. It is similar to the two previous
 lemmas:
@@
 (it must have the correct permission and fit with our stack discipline) and (2)
-require that the adversary callback is global.
+require that the adversary callback be global.
 
@@
 and the adversary is allowed to allocate the activation record on the stack
-(just like we do) so that they can store and recover his old stack pointer after
+(just like we do) so that he can store and recover his old stack pointer after
 \texttt{f4} returns.
@@
 we pass away and static temporary regions (a region that accepts exactly one
-memory) for our local stack as well as the adversary's local stack. The given
+memory) for our local stack as well as the adversaries local stack. The given
 worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
@@
 pointer to return to the adversary, since we have restored validity of any local
-state they might have stashed away.
+state he/she might have stashed away.
 
@@
 
-\textbf{Performance and the requirement for stack clearing} While performance is
-not a goal of this
+\textbf{Performance and the requirement for stack clearing} It is important to
+consider the performance of our proposed stack and return management discipline.
+Generally, the additional security measures described in
+Section~\ref{sec:stack-and-return-pointer} obviously impose an overhead w.r.t.
+an approach that does not provide security. However, they generally consist of a
+few checks or register clearings on every boundary crossing between trusted code
+and adversary, which should impose an acceptable performance overhead. The only
+exception is the requirements that we have for stack clearing, in two
+situations: when returning to the adversary and when invoking an adversary
+callback. As we have explained, it is not sufficient to clear just the part of
+the stack that we have used ourselves, but we need to clear the entire stack,
+except the part that we are using ourselves. This means that on every crossing
+of the boundary between trusted and adversary code, a potentially large region
+of memory needs to be zeroed out entirely. We believe that capability machines
+like CHERI should consider to provide special support for this requirement: a
+highly-optimized instruction for zeroing out a potentially large block of
+memory. We have no experience in processor design, but we imagine this might be
+optimized as a vector instruction or even further, by having the processor keep
+track of a range of memory for which loads should pretend that it has already
+been cleared, by returning zero irrespective of the actual contents of the
+memory.
 
@@
 apply to the ones we get from the adversary (return pointers are allowed to be
-local, callbacks must be global), i.e.\ we do not claim special status and what
+local, callbacks must be global), i.e. we do not claim special status and what
 we propose is modular.

### 2017-02-14 csf-2017/paper.tex

--- 2017-02-14 csf-2017/paper.tex
@@
 \usepackage{multirow,bigdelim}
+
+\ifCLASSOPTIONcompsoc
+\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+\else
+\usepackage[caption=false,font=footnotesize]{subfig}
+\fi
 
@@
 before we do (erasing any data we do not want to pass as (part of) the return
-value) and similarly, erase any part of the stack that we have used (i.e.\
-overwrite it with zeros). Note that we can use the stack as normal: we may push
-values onto it and pop them back off (e.g. for register spilling etc.).
+value) and similarly. However, it is important to carefully consider what needs
+to happen with the stack. Since we may have used the stack for various purposes
+(register spilling, storing local state when invoking other functions etc.), we
+clearly need to clear that data before returning to the adversary, to prevent
+the adversary from gaining access to it.
+
+However, if we are returning from a function that has itself invoked adversary
+code, then even the unused part of the stack may still contain data and
+capabilities left there by them, including local capabilities since the stack is
+store-local. The adversary could take advantage of this by using the unused part
+of the stack to store either stack pointers or return pointers from such
+previous invocations and then use them when they're no longer supposed to. To
+prevent this, before returning, we need to clear (i.e.\ overwrite with zeros) the
+entire part of the stack that the adversary has had access to, not just the
+parts that we have used ourselves. Essentially, this is needed for the same
+reason that we clear the register file: because the adversary should not be able
+to keep hold of local capabilities when control is passed to the trusted code
+and back to the adversary. Clearing the entire unused part of the stack is the
+most problematic aspect of our discipline in terms of efficiency, since it may
+consist of a large part of memory. However, we believe there are ways to
+optimize it that we discuss in Section~\ref{sec:discussion}.
 
@@
   \centering
+
+  \begin{minipage}[t]{4.1cm}
   \begin{lstlisting}
@@
   \end{lstlisting}
-
+  \end{minipage}
+  \begin{minipage}[t]{4.1cm}
   \begin{lstlisting}
@@
   \end{lstlisting}
+  \end{minipage}
   \caption{Two simple programs that rely on local-state encapsulation. \texttt{f1} uses our stack-based calling convention. \texttt{f2} uses a calling not reliant on the stack}

### 2017-02-14 csf-2017/paper.tex
Moved program example g1 and f3 to the same figure.

@@
 \fi
+
+% Include the following to reduce room around display equations
+%\setlength{\abovedisplayskip}{3pt}
+%\setlength{\belowdisplayskip}{3pt}
+%\setlength{\abovedisplayshortskip}{3pt}
+%\setlength{\belowdisplayshortskip}{3pt}
 
@@
 well-bracketed control-flow. To illustrate this, we look at
-two example programs \texttt{\footnotesize{f3}} in \figurename~\ref{fig:prog-f3} and
-\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-g1}.
+two example programs \texttt{\footnotesize{f3}} and
+\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-f3-and-g1}.
 
@@
 lemmas:
+
 \begin{figure}[t]
   \centering
-  \begin{lstlisting}
-f3: push 1
-    fetch $r_1$ $\adv$
-    scall $r_1$($[]$,$[r_1]$)
-    pop $r_2$
-    assert $r_2$ 1
-    push 2
-    scall $r_1$($[]$,$[]$)
-    halt
-\end{lstlisting}
-  \caption{Simple program that relies on control-flow integrity.}
-  \label{fig:prog-f3}
-\end{figure}
-
-\begin{lemma}
-  \label{lem:correctness-f3}
-  Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1}
-  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
-  (\halted,m')$, then $m'(\flag) = 0$.
-\end{lemma}
-
-The final example, in \figurename~\ref{fig:prog-g1}, is a faithful translation
-of a tricky example known from the
-literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
-parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
-generates closures with one variable $x$ set to $0$ in its environment and
-\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
-program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
-sets $x$ to $1$ and calls the callback a second time. When it returns from the
-callback, it asserts that $x$ is $1$ and returns. This example is more
-complicated than the previous ones because it involves a closure invoked by the
-adversary and an adversary callback invoked by us. As explained in
-Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
-to properly check the stack pointer received by our closure from the adversary
-(it must have the correct permission and fit with our stack discipline) and (2)
-require that the adversary callback is global.
-
-\begin{figure}[t]
-  \centering
+  \begin{minipage}[t]{5.05cm}
   \begin{lstlisting}
@@
 \end{lstlisting}
-  \caption{A program that relies on well-bracketedness of
+  \end{minipage}
+  \begin{minipage}[t]{3.4cm}
+  \begin{lstlisting}
+f3: push 1
+    fetch $r_1$ $\adv$
+    scall $r_1$($[]$,$[r_1]$)
+    pop $r_2$
+    assert $r_2$ 1
+    push 2
+    scall $r_1$($[]$,$[]$)
+    halt
+\end{lstlisting}
+  \end{minipage}
+  \caption{ Two programs that rely on well-bracketedness of
     \texttt{scall}s to function correctly. $\var{offset}$ is the
     offset to \texttt{f4}.}
-  \label{fig:prog-g1}
+  \label{fig:prog-f3-and-g1}
 \end{figure}
+
+\begin{lemma}
+  \label{lem:correctness-f3}
+  Making similar assumptions about capabilities and linking as in
+  Lemma~\ref{lem:correctness-f1}
+  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
+  (\halted,m')$, then $m'(\flag) = 0$.
+\end{lemma}
+
+The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}}, is a faithful translation
+of a tricky example known from the
+literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
+parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
+generates closures with one variable $x$ set to $0$ in its environment and
+\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
+program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
+sets $x$ to $1$ and calls the callback a second time. When it returns from the
+callback, it asserts that $x$ is $1$ and returns. This example is more
+complicated than the previous ones because it involves a closure invoked by the
+adversary and an adversary callback invoked by us. As explained in
+Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
+to properly check the stack pointer received by our closure from the adversary
+(it must have the correct permission and fit with our stack discipline) and (2)
+require that the adversary callback is global.
 
@@
 \bibliography{IEEEabrv,references}
+
 \end{document}

### 2017-02-15 csf-2017/paper.tex
inlined updatePcPerm in figure 3

@@
   \begin{align*}
-    \updatePcPerm{w} &=
-                       \begin{cases}
-                         ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
-                         w & \text{\footnotesize{otherwise}}
-                       \end{cases}\\
     \stdUpdatePc{\Phi} &=
@@
                                        \stdUpdatePc{}(\updateReg{r_1}{\\(\permp',\start,\addrend,\addr)})
-                                     \end{multlined}$  &
-                                                        $\memreg(r_2) = \stdcap[\permp]$\footnotesize{ and }\newline$\permp' = \decodePermPair{\memreg(r_2)}$
+                                     \end{multlined}$  & $\memreg(r_2) = \stdcap[\permp]$\footnotesize{ and }$\permp' = \decodePermPair{\memreg(r_2)}$
                                                       \footnotesize{ and }$\permp'\sqsubseteq \permp$\\
@@
     \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$&$\var{newPc} = \updatePcPerm{\memreg(\lv)}$\\
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\perm,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ \footnotesize{ otherwise }$\var{newPc} = \memreg(r)$\\
     \hline
@@
 \subsection{Worlds}
-\dominique{Perhaps we could first try to convey the intuition of a permanent
-  world with a temporary overlay, with global capabilities ignoring the overlay.
-  Perhaps we could add a depiction of the world to Figure 5?}
-
+% \dominique{Perhaps we could first try to convey the intuition of a permanent
+%   world with a temporary overlay, with global capabilities ignoring the overlay.
+%   Perhaps we could add a depiction of the world to Figure 5?}
 %\todo[inline]{Find a good place to describe monotonicity of
@@
 %subsection introduction
-\dominique{I would try to motivate the LR defs by saying that the LR is
-  constructed to express the FTLR and the FTLR says that if a piece of code only
-  has access to values respecting the invariants in a world, then it will
-  respect them itself too.} The logical relation defines semantically what it
-means to be capability safe.\dominique{We have not introduced capability
-  safety..} In large parts, capability safety is defined by when a capability
+% \dominique{I would try to motivate the LR defs by saying that the LR is
+%   constructed to express the FTLR and the FTLR says that if a piece of code only
+%   has access to values respecting the invariants in a world, then it will
+%   respect them itself too.} 
+The logical relation defines semantically what it
+means to be capability safe.
+% \dominique{We have not introduced capability
+%   safety..} 
+In large parts, capability safety is defined by when a capability
 can be safely given to an untrusted piece of code. In the following paragraphs,
@@
 
+\begin{comment}
+\newsavebox{\testit}
+\begin{lrbox}{\testit}
+  \begin{lstlisting}
+g1:  malloc $r_2$ 1
+     store $r_2$ 0
+     move $\pcreg$ $r_3$
+     lea $r_3$ $\var{offset}$
+     crtcls $[(x, r_2)]$ $r_3$
+     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
+     jmp $r_0$
+f4:  reqglob $r_1$
+     prepstk $r_\stk$
+     store $x$ 0
+     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
+     store $x$ 1
+     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
+     load $r_1$ $x$
+     assert $r_1$ 1
+     mclear $r_\stk$
+     rclear $\RegName \setminus \{r_0,\pcreg\}$
+     jmp $r_0$
+\end{lstlisting}
+\end{lrbox}
+
+\newsavebox{\progfthree}
+\begin{lrbox}{\progfthree}
+  \begin{lstlisting}
+f3: push 1
+    fetch $r_1$ $\adv$
+    scall $r_1$($[]$,$[r_1]$)
+    pop $r_2$
+    assert $r_2$ 1
+    push 2
+    scall $r_1$($[]$,$[]$)
+    halt
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}[t]
+  \centering
+  \subfloat{\usebox{\testit}}
+  \subfloat{\usebox{\progfthree}}
+  \caption{test}
+\end{figure}
+\end{comment}
 \begin{figure}[t]

### 2017-02-15 csf-2017/paper.tex
shuffled around restrict equation in figure 3
updated column sizes for figure 3
Moved figure 3 to have it printed on another page
added technical appendix entry MISSING URL!

@@
 \usepackage{listings}
-\usepackage{hyperref}
+\usepackage[draft]{hyperref}
 \usepackage{multicol}
@@
 contiguous C stack and splits it into a central, trusted shadow stack, managed by
-trusted call and return instructions\footnote{These call and return instructions
-  are implemented by CheriBSD and special support from the CHERI processor makes
-  them very fast to invoke.}, and disjoint, per-component
+trusted call and return instructions, and disjoint, per-component
 stacks. To prevent illegal use of stack references,
@@
 For reasons of space, some details and all proofs have been omitted;
-please refer to the technical appendix~\cite{XXX} for those.
+refer to the technical appendix~\cite{technical_appendix} for those.
 
@@
 
-We define the syntax of our capability machine in
-\figurename~\ref{fig:syntax}. We assume an infinite set of addresses
-$\Addrs$ and define machine words as either integers or capabilities
-of the form $((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability
-represents the authority to execute permissions $\perm$ on the memory
-array $[\var{base},\var{end}]$, together with a current address $\addr$ and
-a tag $\gl$ which indicates whether the capability is global or
-local. The capability machine has no notion of pointers by itself, so
-we refer to capabilities as pointers and capabilities
-interchangeably. The available permissions and their ordering are
-depicted in \figurename~\ref{fig:perm-hier}: the permissions include
-null permission ($\noperm$), readonly ($\readonly$), read/write
-($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$)
-permissions. Additionally, there are three special read/write-local
-($\readwritel$), read/write-local/execute ($\rwlx$) and enter
-($\entry$) permissions which we will explain below.
-
-We assume a finite set of register names $\RegName$. We define register files $\reg$ as
-functions mapping register names to words and memories as functions that map
-addresses to words. The state of the entire machine is represented as a
-configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
-and a register file, or a failed or halted state, where the latter keeps hold
-of the final state of memory. 
-% We also define memory segments $\ms$:
-% parts of memory represented as partial functions mapping addresses to words.
-%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
-
-The instruction set of our machine is rather basic. Instructions $i$ include
-relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
-($\move{}{}$, copies words between registers) instructions. Also familiar are
-load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and an arithmetic addition operators ($\plus{}{}{}$ and $\minus{}{}{}$). There are
-three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
-address), $\restricttwo{}{}$ (modifies the permission and local/global flag) and
-$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
-instructions take care that the resulting capability always carries less
-authority than the original (e.g. restrict will only weaken a permission).
-Finally, there is an instruction $\isptr{}{}$ for testing whether a word is a
-capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$,
-$\gete{}{}$ and $\geta{}{}$ for obtaining respectively the permissions,
-local/global flag, base, end and current address of a capability.
-
 \begin{figure*}
@@
   \end{align*}
-
-  \begin{tabular}{|c|p{3.8cm}|p{10.4cm}|}
+  \begin{tabular}{|c|p{4.3cm}|p{9.9cm}|}
     \hline
@@
     \hline
-    $\move{\refreg{r_1}}{r_2}$& $\begin{multlined}
-      \stdUpdatePc{}(\updateReg{r_1}{\\
-          \memreg(r_2)}) \end{multlined}$&\\
+    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)})$&\\
     \hline
@@
     \hline
-    $\restricttwo{\refreg{r_1}}{r_2}$&$\begin{multlined}
-                                       \stdUpdatePc{}(\updateReg{r_1}{\\(\permp',\start,\addrend,\addr)})
-                                     \end{multlined}$  & $\memreg(r_2) = \stdcap[\permp]$\footnotesize{ and }$\permp' = \decodePermPair{\memreg(r_2)}$
-                                                      \footnotesize{ and }$\permp'\sqsubseteq \permp$\\
+    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$\footnotesize{ and }$(\perm',g') = \decodePermPair{\memreg(r_2)}$
+                                                      \footnotesize{ and }$(\perm',g') \sqsubseteq (\perm,g)$\footnotesize{ and } $w =((\perm',g'),\start,\addrend,\addr)$\\
     \hline
@@
   \end{tabular}
-  \caption{An excerpt from the operational semantics}
+  \caption{An excerpt from the operational semantics.}
   \label{fig:op-sem}
 \end{figure*}
+
+We define the syntax of our capability machine in
+\figurename~\ref{fig:syntax}. We assume an infinite set of addresses
+$\Addrs$ and define machine words as either integers or capabilities
+of the form $((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability
+represents the authority to execute permissions $\perm$ on the memory
+array $[\var{base},\var{end}]$, together with a current address $\addr$ and
+a tag $\gl$ which indicates whether the capability is global or
+local. The capability machine has no notion of pointers by itself, so
+we refer to capabilities as pointers and capabilities
+interchangeably. The available permissions and their ordering are
+depicted in \figurename~\ref{fig:perm-hier}: the permissions include
+null permission ($\noperm$), readonly ($\readonly$), read/write
+($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$)
+permissions. Additionally, there are three special read/write-local
+($\readwritel$), read/write-local/execute ($\rwlx$) and enter
+($\entry$) permissions which we will explain below.
+
+We assume a set of register names $\RegName$. We define register files $\reg$ as
+functions mapping register names to words and memories as functions that map
+addresses to words. The state of the entire machine is represented as a
+configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
+and a register file, or a failed or halted state, where the latter keeps hold
+of the final state of memory. 
+% We also define memory segments $\ms$:
+% parts of memory represented as partial functions mapping addresses to words.
+%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
+
+The instruction set of our machine is rather basic. Instructions $i$ include
+relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
+($\move{}{}$, copies words between registers) instructions. Also familiar are
+load and store instructions for reading from and writing to memory ($\load{}{}$
+and $\store{}{}$) and an arithmetic addition operators ($\plus{}{}{}$ and $\minus{}{}{}$). There are
+three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
+address), $\restricttwo{}{}$ (modifies the permission and local/global flag) and
+$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
+instructions take care that the resulting capability always carries less
+authority than the original (e.g. restrict will only weaken a permission).
+Finally, there is an instruction $\isptr{}{}$ for testing whether a word is a
+capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$,
+$\gete{}{}$ and $\geta{}{}$ for obtaining respectively the permissions,
+local/global flag, base, end and current address of a capability.
 
@@
 the model linking is described to the degree necessary, in the next
-section, but we refer to the technical appendix~\ref{XXX} for further
+section, but we refer to the technical appendix~\cite{technical_appendix} for further
 details.
@@
 starting with first-order interfaces and moving onto a higher-order setting
-after that. After that, we define a number of reusable macro-instructions that
+next. After that, we define a number of reusable macro-instructions that
 can be used to conveniently apply the proposed discipline in subsequent
@@
 purpose. We stress that the stack is not built-in, but merely an
-abstraction we put on a piece of the memorey. This stack pointer must be local and it must carry $\rwlx$ permission. It is crucial for
+abstraction we put on this piece of the memorey. This stack pointer must be local and it must carry $\rwlx$ permission. It is crucial for
 our approach that the stack is the only part of memory for which there will ever
@@
 (register spilling, storing local state when invoking other functions etc.), we
-clearly need to clear that data before returning to the adversary, to prevent
-the adversary from gaining access to it.
+clearly need to erase that data before returning to the adversary, to prevent
+the adversary access to it.
 
@@
 In Section~\ref{sec:examples}, we use the above macros and some
-more. The technical appendix~\ref{XXX} contains detailed descriptions
+more. The technical appendix~\cite{technical_appendix} contains detailed descriptions
 of all the macros.
@@
 Some definitions are omitted and explained when needed. The precise definitions
-can be found in the technical appendix~\cite{XXX}.
+can be found in the technical appendix~\cite{technical_appendix}.
 
@@
 The interested reader can find all the technical details in the
-technical appendix~\cite{XXX}.
+technical appendix~\cite{technical_appendix}.
 
@@
 \label{sec:conclusion}
-
+\todo[inline]{Insert the correct URL for the technical report in the bibtex file!}
 \section*{Acknowledgements}

### 2017-02-15 local-capability-machine.tex
comment about call macro
slight change in f1 lemma
added minus instruction
removed comments and time stamp, added title

@@
 \newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+
+%Disable comments (uncomment the following)
+\renewcommand\lau[1]{}
+\renewcommand\dominique[1]{}
+\renewcommand\lars[1]{}
 
@@
 \newcommand\bnfdef{\mathrel{::=}}
+\title{Reasoning about a Capability Machine with Local Capabilities\\
+ Provably Safe Stack and Return Pointer Management (without OS Support)\\
+ Technical Appendix Including Proofs and Details}
 
 \begin{document}
-\begin{flushright}
-  \today
-\end{flushright}
-
+\maketitle
 \tableofcontents
@@
 \begin{align*}
-  \executeAllowed{\perm} &=
-                           \begin{cases}
-                             \true & \text{if } \perm \in \{ \rwx, \rwlx, \exec, \entry \} \\
-                             \false & \text{otherwise}
-                           \end{cases} \\
+  % \executeAllowed{\perm} &=
+  %                          \begin{cases}
+  %                            \true & \text{if } \perm \in \{ \rwx, \rwlx, \exec, \entry \} \\
+  %                            \false & \text{otherwise}
+  %                          \end{cases} \\
   \readAllowed{\perm} &=
@@
                                                             \stdUpdatePc{\updateReg{r_1}{n_1+n_2}} &
+                                                            \arraycolsep=0pt
+                                                            \begin{array}[t]{l}
+                                                              \text{if for $i \in \{1,2\}$}\\
+                                                              \quad\text{$n_i = \rv_i$ or $n_i = \memreg(\rv_i)$}\\
+                                                              \quad\text{and in either case $n_i \in \ints$}
+                                                            \end{array}\\
+                                                            \failed & \text{otherwise}
+                                                          \end{cases}\\
+  \sem{\minus{\refreg{r_1}}{\rv_1}{\rv_2}}               & =
+                                                          \begin{cases}
+                                                            \stdUpdatePc{\updateReg{r_1}{n_1-n_2}} &
                                                             \arraycolsep=0pt
@@
               \item \texttt{call} will fail if we have local capabilities in one of the registers of the ``private'' register list as it relies on a capability returned by malloc which will not be permit-write-local.
+This severely limits how \texttt{scall} can be used and it provides very little in terms of control-flow integrety when nested.
 Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
@@
                 \begin{itemize}
+                \item Push ``private'' registers to the stack.
                 \item Push the restore code to the stack.
-                \item Push ``private'' registers to the stack.
                 \item Push return address capability 
@@
                 \begin{description}
+                \item [Push ``private'' registers to the stack]
+                  Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
                 \item [Push the restore code to the stack]
@@
 The caller code does the rest of the restoration.
-                \item [Push ``private'' registers to the stack]
-                  Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
                 \item [Push return address capability]
@@
                   \begin{itemize}
+                  \item Pop the restore code of the stack
                   \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
-                  \item Pop the restore code of the stack
                   \end{itemize}\
@@
   \[
-    (\reg_0\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
+    (\reg\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
   \]
@@
   \[
-    (\reg_0\update{\pcreg}{c_{f1}},m) \step[i] (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}'}\update{r_1}{1},m)
+    (\reg\update{\pcreg}{c_{f1}},m) \step[i] (\reg_0\update{\pcreg}{c_\malloc}\update{r_0}{c_{f1}'}\update{r_1}{1},m)
   \]

### 2017-02-15 csf-2017/paper.tex
added a couple of words to prevent line break in reference causing hyperlink to break
Some work on discussion and RW.

--- 2017-02-15 csf-2017/paper.tex
@@
   \end{tabular}
-  \caption{An excerpt from the operational semantics}
+  \caption{An excerpt from the operational semantics.}
   \label{fig:op-sem}
 \end{figure*}
+
+We define the syntax of our capability machine in
+\figurename~\ref{fig:syntax}. We assume an infinite set of addresses
+$\Addrs$ and define machine words as either integers or capabilities
+of the form $((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability
+represents the authority to execute permissions $\perm$ on the memory
+array $[\var{base},\var{end}]$, together with a current address $\addr$ and
+a tag $\gl$ which indicates whether the capability is global or
+local. The capability machine has no notion of pointers by itself, so
+we refer to capabilities as pointers and capabilities
+interchangeably. The available permissions and their ordering are
+depicted in \figurename~\ref{fig:perm-hier}: the permissions include
+null permission ($\noperm$), readonly ($\readonly$), read/write
+($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$)
+permissions. Additionally, there are three special read/write-local
+($\readwritel$), read/write-local/execute ($\rwlx$) and enter
+($\entry$) permissions which we will explain below.
+
+We assume a finite set of register names $\RegName$. We define register files $\reg$ as
+functions mapping register names to words and memories as functions that map
+addresses to words. The state of the entire machine is represented as a
+configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
+and a register file, or a failed or halted state, where the latter keeps hold
+of the final state of memory. 
+% We also define memory segments $\ms$:
+% parts of memory represented as partial functions mapping addresses to words.
+%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
+
+The instruction set of our machine is rather basic. Instructions $i$ include
+relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
+($\move{}{}$, copies words between registers) instructions. Also familiar are
+load and store instructions for reading from and writing to memory ($\load{}{}$
+and $\store{}{}$) and an arithmetic addition operators ($\plus{}{}{}$ and $\minus{}{}{}$). There are
+three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
+address), $\restricttwo{}{}$ (modifies the permission and local/global flag) and
+$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
+instructions take care that the resulting capability always carries less
+authority than the original (e.g. restrict will only weaken a permission).
+Finally, there is an instruction $\isptr{}{}$ for testing whether a word is a
+capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$,
+$\gete{}{}$ and $\geta{}{}$ for obtaining respectively the permissions,
+local/global flag, base, end and current address of a capability.
 
@@
 
+\subsection{Stack and return pointer discipline}
 \textbf{Formulating control flow integrity} The attentive reader may notice
@@
 of the boundary between trusted and adversary code, a potentially large region
-of memory needs to be zeroed out entirely. We believe that capability machines
-like CHERI should consider to provide special support for this requirement: a
-highly-optimized instruction for zeroing out a potentially large block of
-memory. We have no experience in processor design, but we imagine this might be
-optimized as a vector instruction or even further, by having the processor keep
-track of a range of memory for which loads should pretend that it has already
-been cleared, by returning zero irrespective of the actual contents of the
-memory.
-
-\textbf{Modularity} Our return pointers and callbacks satisfy the checks we
-apply to the ones we get from the adversary (return pointers are allowed to be
-local, callbacks must be global), i.e.\ we do not claim special status and what
-we propose is modular.
+of memory needs to be zeroed out entirely. We believe that this in fact is a
+common requirement for typical usage scenario's of local capabilities and
+capability machines like CHERI should consider to provide special support for
+this requirement, in the form of a highly-optimized instruction for zeroing out
+a potentially large block of memory. We have no experience in processor design,
+but we imagine this might be optimized as a vector instruction or even further,
+by having the processor keep track of a range of memory for which loads should
+pretend that it has already been cleared, by returning zero irrespective of the
+actual contents of the memory.
+
+\textbf{Modularity} An important aspect of our stack and return pointer
+discipline is that it is modular. By that, we mean that the requirements we have
+on callbacks, return pointers etc. provided to trusted code by an adversary are
+also satisfied by callbacks, return pointers etc. that we construct ourselves.
+For example, our return pointers are necessarily local capabilities because they
+must point to memory where we can store the old stack pointer, but the
+adversary's return pointers are also allowed to be local. Adversary callbacks on
+the other hand are required to be global but the callbacks we construct
+ourselves are also allocated on the heap and global. In other words, we do not
+assume that our code has some sort of special status and it is possible for the
+adversary to protect themselves against us in the same way as we protect against
+them.
+
+\subsection{Logical relation}
+
+\textbf{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in
+Figure~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an
+existing pattern called biorthogonality. Biorthogonality is a pattern for
+defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998}
+in terms of an observation relation of safe configurations (like we do). The
+idea is to define safe evaluation contexts as the set of contexts that produce
+safe observations when plugging safe values and define safe terms as the set of
+terms that can be plugged into safe evaluation contexts to produce safe values.
+This is an alternative to more direct definitions where safe terms are defined
+as terms that evaluate to safe values. An advantage of biorthogonality is that
+it scales better to languages with control effects like call/cc. Our definitions
+can be seen as a variant of biorthogonality, where we take only a single
+orthogonal closure: we do not define safe evaluation contexts but immediately
+define safe terms as those that produce safe observations when plugged with safe
+values. This is natural because we model arbitrary assembly code that does not
+necessarily respect a particular calling convention. As a result, return
+pointers are in principle values like all others and there is no reason to treat
+them specially in the logical relation.
+
+It is interesting to note that \citet{Hur:2011:KLR:1926385.1926402} also build a
+step-indexed, Kripke, biorthogonal logical relation for an assembly language
+(for reasoning about correct compilation from ML to assembly), but because they
+only model non-adversarial code that treats return pointers specially according
+to a particular calling convention, they can use standard biorthogonality rather
+than a single orthogonal closure like us.
+
+\textbf{Public/private future worlds} A novel aspect of our logical relation is
+how we model the temporary nature of local capabilities using public/private
+future worlds. The main insight is that the temporary, revokable nature of these
+capabilities can be seen as a generalisation of the syntactically-enforced
+unstoreable status of evaluation contexts in lambda calculi without control
+effects (of which well-bracketed control flow is a consequence). To reason about
+code that relies on this (particularly, the original awkward example),
+\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
+contexts using Kripke worlds with public and private future world relations.
+Essentially, they allow relatedness of evaluation contexts to be monotone with
+respect to a weaker future world relation (public) than relatedness of values,
+formalising the idea that it is safe to make temporary internal state
+modifications (private world transititions, that invalidate the continuation,
+but not other values) while an expression is performing internal steps, as long
+as the code returns to a stable state (i.e. transitions to a public future world
+of the original) before returning (i.e. invoking the continuation). We
+generalise this idea to reason about local capabilities: validity of local
+capabilities is allowed to be monotone with respect to a weaker future-world
+relation than other values, which we can exploit to distinguish between state
+changes that are always safe (public future worlds) and changes that are only
+valid if we clear all local capabilities (private future worlds). Our future
+world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
+the awkward example uses exactly the same state transition system than theirs),
+but they turn up in an entirely different place in the logical relation: rather
+than using public future worlds for the special syntactic category of evaluation
+contexts, they are used in the value relation depending on the depending on the
+locality of the capability at hand. Additionally, our worlds are a bit more
+complex because, to allow local memory capabilities and store-local
+capabilities, they can contain (revokable) temporary regions that are only
+monotonous w.r.t. public future worlds, while DNB's worlds are entirely
+permanent.
 
 \textbf{Local capabilities in high-level languages} It is worth mentioning that
-local capabilities are quite similar to features that have been proposed for
-high-level languages. Specifically, \citet{osvald_gentrification_2016}'s
-second-class or local values in Scala are remarkably similar: they are a kind of
-values that can be provided to other code for immediate use without allowing it
-to store the value in a closure or reference for later use. We believe reasoning
-about such values will require very similar reasoning techniques than the ones
-we have used for local capabilities.
+local capabilities are quite similar to a feature proposed for the high-level
+language Scala: \citet{osvald_gentrification_2016}'s second-class or local
+values. They are a kind of values that can be provided to other code for
+immediate use without allowing it to be stored in a closure or reference for
+later use. We believe reasoning about such values will require very similar
+reasoning techniques than the ones we have used for local capabilities.
 
@@
 externally observable side-effects (like console output or memory access
-traces), we do not require their notion of effect parametricity.
-
-The idea of applying SKLRs to an assembly language was applied before by
-\citet{Hur:2011:KLR:1926385.1926402}, who constructed a binary SKLR for
-formulating and proving correctness of an ML-to-assembly compiler. Their
-condition on the assembly code are quite loose (for example, they allow
-hand-written and self-modifying assembly code), but they do assume code that is
-non-adversarial and respects the calling convention of the high-level language
-(for example, the code does not store return pointers and invoke them after
-their intended lifetime). As a result, they can treat return pointers as
-evaluation contexts (continuations) in a lambda calculus and use biorthogonality
-to define the LR~\cite{pitts_operational_1998}. In our setting, there is no
-guarantee that potentially adversarial code will respect the calling convention,
-so, as described in Section~\ref{sec:logical-relation}, we do not use
-biorthogonal closure but just model return pointers as continuations using a
-single orthogonal closure.
-
-Another novel aspect of our logical relation is how we model the temporary
-nature of local capabilities. The main insight is that the temporary, revokable
-nature of these capabilities can be seen as a generalisation of the
-syntactically special, unstoreable status of evaluation contexts in lambda
-calculi (without control effects). Closely related to this is a study by
-\citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and control
-effects on local relational reasoning. DNB formally capture the special status
-of evaluation contexts using Kripke worlds with public and private future world
-relations. Our future world relations are inspired by theirs and they are at
-times remarkably similar. For example, the state-transitition system in the
-proof of our assembly version of the awkward example, is exactly the same as the
-one used by DNB for the original awkward example. However, in our logical
-relation, public and private future worlds turn up in an entirely different
-place in the logical relation: rather than using public future worlds for the
-special syntactic category of evaluation contexts, they are used in the value
-relation depending on the depending on the locality of the capability at hand.
-Additionally, our worlds are a bit more complex because, to allow store-local
-capabilities, they can have temporary regions that are only monotonous w.r.t.
-public future worlds, while DNB's worlds are entirely permanent.
-
-%/comparison with related Dreyer-Neis-Birkedal and with Hur-Dreyer}
-%\todo[inline]{Add discussion/comparison to other LRs}
-% Comparisson of LR to other work:
-% - Single othorgonal closure
-%   o all capabilities are possible continuations, not just the one we
-%   intend to be it.
-% - No termination guarantees in observation relation
-% - Private/public transitions used like in Dreyer-Neis-Birkedal, but
-%   also to  handle the difference in what local and global
-%   capabilities can handle.
-
-% As already mentioned, our the state transition systems used in our
-% logical relation are inspired by earlier work on logical relations for
-% high-level
-% languages~\cite{Ahmed:popl09, Dreyer:jfp12,Devriese:2016ObjCap}.
-% The idea of public and private transitions comes from 
-% Dreyer et. al.~\cite{Dreyer:jfp12}, who gave a logical relations model
-% of a high-level lambda calculus. In \emph{loc. cit.}
-% public transitions were used to 
-% capture the \emph{extensional} behavior of a program expression
-% (and a context), whereas private transitions were used to capture
-% the internal behaviour of a computation using local state.
-% Technically, the logical relation of Dreyer et. al.\ was
-% formulated using biorthogonal closure, and the future world
-% relation for continuations was required to be a public future world
-% relation. In other words, the choice between private or public future
-% rested on the ``structure'' of the programming language. 
-% In contrast, 
-% for our low-level, unstructured, assembly language, we
-% we instead decide on whether to use a private or public future 
-% world depending on whether the capability is global or local. 
-% Another interesting technical difference is that in our low-level 
-% language we do not have a fixed calling convention and any register
-% can serve as a continuation, and therefore we do not use biorthogonal
-% closure but just model return pointers as continuations using a single
-% orthogonal closure.
-
-% \lars{don't know yet whether we need something like the following paragraph:}
-% Even though 
-% Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also
-% considered a logical
-% relation involving an assembly language, they still used
-% biorthogonal closure but that was because their logical relation
-% was a relation between a high-level language and an assembly
-% language,  and thus the assembly language in the relation essentially
-% inherited structure from the high-level language.
+traces), we do not require their notion of effect parametricity. Our logical
+relation uses several ideas from previous work, like Kripke worlds with regions
+containing state transition systems~\citep{Ahmed:popl09}, public/private future
+worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion),
+biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
 

### 2017-02-15 csf-2017/paper.tex

@@
 
-We assume a finite set of register names $\RegName$. We define register files $\reg$ as
+We assume a set of register names $\RegName$. We define register files $\reg$ as
 functions mapping register names to words and memories as functions that map
@@
 \emph{local capabilities}, modeled after the ones in
-CHERI~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
+the CHERI processor~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
 they can only be kept in registers, i.e.\ they cannot be stored to memory. This

### 2017-02-15 csf-2017/paper.tex
shuffled around stuff in attempt to make tex have less white space on ceratin pages.

@@
 \textbf{Reusable macro instructions} 
-We define a number of reusable macros. Some of them capture the stack
-management discipline and others make the examples in
-Section~\ref{sec:examples} readable. The macros are:
+We define a number of reusable macros. The macros capture the stack
+management discipline and other conveniences. The macros are:
 \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
@@
 
-In Section~\ref{sec:examples}, we use the above macros and some
-more. The technical appendix~\cite{technical_appendix} contains detailed descriptions
+The technical appendix~\cite{technical_appendix} contains detailed descriptions
 of all the macros.
@@
 % \lau{this may not be an important enough point to include it.}
-\begin{figure}
-  \centering
-  \includegraphics{w11}
-  \caption{Illustration of the relation between local and global capabilities and temporary and permanent regions.
-The table represents a memory and the colored fields are regions governing parts of the memory.
-Roughly speaking, global capabilities should never depend on memory governed by temporary regions.}
-  \label{fig:cap-world}
-\end{figure}
 A world is a finite map from region names, modeled as natural numbers, to regions.
@@
   preordered \cofes{}, and there exists an isomorphism $\xi$ such that
-  {\small
   \begin{align*}
-      \xi : \Wor & \cong \blater (\nats \finparfun \Regions)\\
+      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions) \\
       \Regions & = \{\revoked\} \uplus \\
@@
     \end{align*}
-  } and for $W, W' \in \Wor$
+  and for $W, W' \in \Wor$
   \begin{align*}
@@
 the worlds:
-\begin{align*}
-  \Worlds & = \RegionNames \finparfun \Regions\\
- \RegionNames & = \nats
-\end{align*}
+\[
+  \Worlds = \nats \finparfun \Regions
+\]
+
+\begin{figure}
+  \centering
+  \includegraphics{w11}
+  \caption{Illustration of the relation between local and global capabilities and temporary and permanent regions.
+The table represents a memory and the colored fields are regions governing parts of the memory.
+Roughly speaking, global capabilities should never depend on memory governed by temporary regions.}
+  \label{fig:cap-world}
+\end{figure}
 

### 2017-02-15 csf-2017/paper.tex
A (final?) missing piece of the RW.

--- 2017-02-15 csf-2017/paper.tex
@@
 
+In this section, we summarise how our work relates to previous work. We do not
+repeat the work already discussed in Section~\ref{sec:discussion}.
+
 Capability machines originate with Dennis and Van
@@
 
-Other papers that enforce well-bracketed control flow at low level (using a
-trusted stack manager)
-\begin{itemize}
-\item \cite{patrignani_modular_2016-1}
-\item \cite{juglaret_beyond_2016-1}
-\item other stuff?
-\end{itemize}
-CFI: 
-\begin{itemize}
-\item \cite{budiu_architectural_2006,abadi_control-flow_2005}
-\end{itemize}
+There are plenty of other papers that enforce well-bracketed control flow at a
+low level, but most are restricted to preventing particular types of attacks and
+enforce only partial correctness of control flow. This includes particularly the
+line of work on \emph{control flow integrity}~\citep{abadi_control-flow_2005}.
+This work uses a quite different attacker model than us: they assume an attacker
+that is not able to execute code, but can overwrite arbitrary data at any time
+during execution (as a model for buffer overflows). By checking the address of
+every indirect jump and using memory access control to prevent overwriting code,
+this work enforces what they call control flow integrity, formalised as the
+property that every jump will follow a legal path in the control flow graph.
+As discussed in Section~\ref{sec:discussion}, this correctness property does not
+take into account temporal properties and seems hard to use for reasoning.
+
+More closely related to our work are papers that use a trusted stack manager and
+some form of memory isolation to enforce control flow correctness as part of a
+secure compilation
+result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
+differs from theirs in that we use a different form of low-level security
+primitive (a capability machine with local capabilities rather than a machine
+with a primitive notion of compartments) and we do not use a trusted stack
+manager component, but a decentralised calling convention based on local
+capabilities. Also, they both prove a secure compilation result from a
+high-level language, which clearly implies a general form of control flow
+correctness, while we define a sound logical relation that can be used to reason
+about specific programs that rely on well-bracketed control flow.
 

### 2017-02-15 csf-2017/paper.tex
Rename "discussion and future work" section to just discussion and save a line somewhere.

@@
 
-\section{Discussion, future work}
+\section{Discussion}
 \label{sec:discussion}
@@
 manager component, but a decentralised calling convention based on local
-capabilities. Also, they both prove a secure compilation result from a
-high-level language, which clearly implies a general form of control flow
-correctness, while we define a sound logical relation that can be used to reason
-about specific programs that rely on well-bracketed control flow.
+capabilities. Also, both prove a secure compilation result from a high-level
+language, which clearly implies a general form of control flow correctness,
+while we define a sound logical relation that can be used to reason about
+specific programs that rely on well-bracketed control flow.
 

### 2017-02-15 csf-2017/paper.tex
Refer to Akram's master thesis.

@@
 
-\begin{itemize}
-\item \cite{akram_el-korashy_formal_2016}
-\end{itemize}
+El-Korashy also defined a formal model of a capability machine, namely CHERI,
+and uses it to prove a compartmentalization
+result~\citep{akram_el-korashy_formal_2016} (not implying control flow
+correctness). He also proves soundness of control-flow integrity (see above)
+adapted to the capability machine, a result that seems orthogonal to the fact
+that the machine uses capabilities.
 

### 2017-02-15 csf-2017/paper.tex
Change some \cite's to \citep's.

@@
 verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a native
-stack (e.g.,~the JVM~\cite{lindholm_java_2014}) or
-randomization~\cite{forrest_building_1997}. The latter techniques essentially
+stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
+randomization~\citep{forrest_building_1997}. The latter techniques essentially
 emulate protection techniques on existing hardware, at the cost of performance,
@@
 challenges like backwards compatibility and realistic OS
-support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
+support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
 machines tag every word (in the register file and in memory) to

### 2017-02-15 csf-2017/paper.tex
removed acknowledgement and authors
Rephrase control flow integrity to control flow correctness.

@@
 
-\author{%
-\IEEEauthorblockN{Lau~Skorstengaard}
-\IEEEauthorblockA{Aarhus University, Denmark\\
-Email: lau@cs.au.dk} \and
-\IEEEauthorblockN{Dominique~Devriese}
-  \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
-Email: dominique.devriese@cs.kuleuven.be} \and
-\IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
-Email: birkedal@cs.au.dk}}
-
+% \author{%
+% \IEEEauthorblockN{Lau~Skorstengaard}
+% \IEEEauthorblockA{Aarhus University, Denmark\\
+% Email: lau@cs.au.dk} \and
+% \IEEEauthorblockN{Dominique~Devriese}
+%   \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
+% Email: dominique.devriese@cs.kuleuven.be} \and
+% \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
+% Email: birkedal@cs.au.dk}}
+\author{}
 \maketitle
@@
   which makes them an interesting target for compilation schemes that aim to
-  provably enforce control-flow integrity, encapsulation of local
+  provably enforce control-flow correctness, encapsulation of local
   state, and such properties.
@@
   and propose a novel stack management discipline for enforcing
-  control-flow integrity and encapsulation of local state on it. To prove
+  control-flow correctness and encapsulation of local state on it. To prove
   these properties, we provide a logical relation that semantically
@@
 Compromising software security is often based on attacks that break programming
-language properties relied upon by software authors, such as control flow
-integrity, local state encapsulation, etc. Commodity processors offer little
+language properties relied upon by software authors, such as control-flow
+correctness, local state encapsulation, etc. Commodity processors offer little
 support for defending against such attacks: they offer security primitives with
 only coarse-grained memory protection and limited compartmentalization
-scalability. As a result, defenses against attacks on control flow integrity and
-local state encapsulation are either limited to only certain common forms of
+scalability. As a result, defenses against attacks on control-flow correctness
+and local state encapsulation are either limited to only certain common forms of
 attacks (leading to an attack-defense arms race) or rely on techniques like
@@
 (e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
-verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a native
-stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
+verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a
+native stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
 randomization~\citep{forrest_building_1997}. The latter techniques essentially
@@
 enforcement of properties like local state encapsulation and
-control-flow integrity. Potentially, they will enable new compilation
+control-flow correctness. Potentially, they will enable new compilation
 schemes that enforce such properties in a way that is efficient but
@@
 
-A similar problem is the enforcement of control-flow integrity on capability
-machines. An interesting approach is taken in
-CheriBSD~\citep{Watson2015Cheri}: It drops the standard
-contiguous C stack and splits it into a central, trusted shadow stack, managed by
-trusted call and return instructions, and disjoint, per-component
-stacks. To prevent illegal use of stack references,
-the approach relies on \emph{local capabilities}, a type of capabilities offered
-by CHERI that can be passed over a function invocation \emph{temporarily}, for
-the duration of the invocation, and revoked afterwards. However, a lot of
-details remain to be investigated (e.g., how to combine a trusted shadow stack with
+A similar problem is the enforcement of control-flow correctness on capability
+machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
+It drops the standard contiguous C stack and splits it into a central, trusted
+shadow stack, managed by trusted call and return instructions, and disjoint,
+per-component stacks. To prevent illegal use of stack references, the approach
+relies on \emph{local capabilities}, a type of capabilities offered by CHERI
+that can be passed over a function invocation \emph{temporarily}, for the
+duration of the invocation, and revoked afterwards. However, a lot of details
+remain to be investigated (e.g., how to combine a trusted shadow stack with
 (potentially) cross-domain function pointers in a scalable way?) and there is no
-argument that this enforcement of control-flow integrity is watertight.
+argument that this enforcement of control-flow correctness is watertight.
 
 In this paper, we make progress towards the goal of efficient and provably
-watertight enforcement of encapsulation and control-flow integrity on capability
-machines.  Specifically, we make the following contributions:
+watertight enforcement of encapsulation and control-flow correctness on
+capability machines. Specifically, we make the following contributions:
 \begin{itemize}
@@
 \item We define a novel stack management discipline for enforcing control-flow
-  integrity and encapsulation of local state on the stack in Section~\ref{sec:stack-and-return-pointer}. The discipline relies
-  solely on local capabilities (used to represent stack and return pointers) and
-  does not require OS support in the form of a trusted shadow stack or special
-  call and return instructions. It also does not require separate per-component
-  stacks and instead works with a single, fairly standard, contiguous stack. It
+  correctness and encapsulation of local state on the stack in
+  Section~\ref{sec:stack-and-return-pointer}. The discipline relies solely on
+  local capabilities (used to represent stack and return pointers) and does not
+  require OS support in the form of a trusted shadow stack or special call and
+  return instructions. It also does not require separate per-component stacks
+  and instead works with a single, fairly standard, contiguous stack. It
   supports higher-order cross-component calls (e.g., (potentially)
-  cross-component function pointers) and can be efficient assuming
-  only
-  one additional piece of processor support: an instruction for efficiently clearing
+  cross-component function pointers) and can be efficient assuming only one
+  additional piece of processor support: an instruction for efficiently clearing
   a range of memory.
@@
   examples, specifically constructed to demonstrate local state encapsulation
-  and control-flow integrity guarantees in the presence of cross-component
+  and control-flow correctness guarantees in the presence of cross-component
   function pointers. The examples demonstrate both the power of our formulation
@@
 capabilities on a capability machine can be used in a stack and return pointer
-management scheme that enforces control flow integrity in a way that is provably
+management scheme that enforces control-flow correctness in a way that is provably
 watertight, potentially efficient, does not rely on a trusted central stack
@@
 safety allows us to prove local state encapsulation and control-flow
-integrity properties for challenging program examples. The security
+correctness properties for challenging program examples. The security
 measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
@@
 \subsection{Stack and return pointer discipline}
-\textbf{Formulating control flow integrity} The attentive reader may notice
+\textbf{Formulating control flow correctness} The attentive reader may notice
 that, while we claim that our stack and return pointer discipline enforces
-control flow integrity, we do not prove a general theorem to show that this is
+control-flow correctness, we do not prove a general theorem to show that this is
 the case. The reason for this is that it is not clear what such a theorem should
-look like. Formulations in terms of a control flow graph, like the one by
+look like. Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
-like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.\footnote{It is a
-  bit unfortunate that the general term ``control flow integrity'' is used for a
-  property that only captures partial correctness of control flow.} In fact, our
-examples show that our logical relation imply a form of control flow integrity
+like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on. In fact, our
+examples show that our logical relation imply a form of control-flow correctness
 that is stronger than such formulations, although this is not made very
@@
 More closely related to our work are papers that use a trusted stack manager and
-some form of memory isolation to enforce control flow correctness as part of a
+some form of memory isolation to enforce control-flow correctness as part of a
 secure compilation
@@
 capabilities. Also, both prove a secure compilation result from a high-level
-language, which clearly implies a general form of control flow correctness,
+language, which clearly implies a general form of control-flow correctness,
 while we define a sound logical relation that can be used to reason about
@@
 and uses it to prove a compartmentalization
-result~\citep{akram_el-korashy_formal_2016} (not implying control flow
-correctness). He also proves soundness of control-flow integrity (see above)
+result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
+correctness). He also proves soundness of control flow integrity (see above)
 adapted to the capability machine, a result that seems orthogonal to the fact
@@
 
-\section{Conclusion}
+%\section{Conclusion}
 \label{sec:conclusion}
 \todo[inline]{Insert the correct URL for the technical report in the bibtex file!}
+\begin{comment}
 \section*{Acknowledgements}
@@
 Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
-
+\end{comment}
 %\bibliographystyle{IEEEtran}

### 2017-02-15 csf-2017/paper.tex
Replace the term "stack and return pointer discipline" by "calling convention" and start shortening.

@@
 \begin{abstract}  
-  Capability machines provide security guarantees at machine level
-  which makes them an interesting target for compilation schemes that aim to
-  provably enforce control-flow correctness, encapsulation of local
-  state, and such properties.
-  We provide a formalisation of a representative capability machine
-  and propose a novel stack management discipline for enforcing
-  control-flow correctness and encapsulation of local state on it. To prove
-  these properties, we provide a logical relation that semantically
-  captures capability safety. The capability safety defined by the
-  logical relation goes beyond the
-  proposed stack management discipline and can be used to reason about
-  any capability safe program.
+  Capability machines provide security guarantees at machine level which makes
+  them an interesting target for compilation schemes that aim to provably
+  enforce properties like control-flow correctness, encapsulation of local
+  state. We provide a formalisation of a representative capability machine and
+  propose a novel calling convention for enforcing control-flow correctness and
+  encapsulation of local state on it. To prove these properties, we provide a
+  logical relation that semantically captures the guarantees provided by the
+  hardware (a form of capability safety). These results are not tied to our
+  calling convention and can be used to reason about arbitrary programs.
 \end{abstract}
@@
 \begin{itemize}
-\item In Section~\ref{sec:capab-mach-with}, we formalise a simple but representative capability machine featuring
-  local capabilities much like CHERI's, and define its operational semantics.
-\item We define a novel stack management discipline for enforcing control-flow
-  correctness and encapsulation of local state on the stack in
-  Section~\ref{sec:stack-and-return-pointer}. The discipline relies solely on
-  local capabilities (used to represent stack and return pointers) and does not
-  require OS support in the form of a trusted shadow stack or special call and
-  return instructions. It also does not require separate per-component stacks
-  and instead works with a single, fairly standard, contiguous stack. It
-  supports higher-order cross-component calls (e.g., (potentially)
-  cross-component function pointers) and can be efficient assuming only one
-  additional piece of processor support: an instruction for efficiently clearing
-  a range of memory.
-\item We provide a theorem that is (by far) the most general
-  and powerful formulation of the guarantees offered by a capability machine.
-  The theorem also expresses the specific guarantees offered by the machine for
-  local capabilities. It is very general and not tied to a specific ABI or a
-  specific way of using the system's capabilities.
-\item The above theorem is formulated in terms of a unary step-indexed
-  logical relation with recursively defined Kripke worlds. It is
-  inspired by one used in \cite{Devriese:2016ObjCap} for
-  capability safety in a JavaScript-like lambda calculus, but has two
-  novel technical characteristics. The natural CPS style of assembly
-  code with return pointers as continuations is modeled using a
-  \emph{single} orthogonal closure, rather than the biorthogonal
-  closure that has been previously used for lambda calculi. Perhaps
-  more surprisingly, we show that the special nature of local
-  capabilities can be elegantly expressed using a variant of an idea
-  that has been used by Dreyer et. al.~\cite{Dreyer:jfp12} to express
-  well-bracketedness guarantees in a lambda calculus: Kripke worlds
-  with both a public and a private future world relation. Interestingly,
-  the two future world relations generally work very similarly to
-  Dreyer et al.'s, except that they turn up in quite a different place
-  in the logical relation. The logical relation and theorem is presented in Section~\ref{sec:logical-relation}.
-\item We demonstrate the power of our results in Section~\ref{sec:examples} by applying them to challenging
-  examples, specifically constructed to demonstrate local state encapsulation
-  and control-flow correctness guarantees in the presence of cross-component
-  function pointers. The examples demonstrate both the power of our formulation
-  of capability safety and our stack management discipline.
+\item Formalise a simple but representative capability machine featuring local
+  capabilities, and its operational semantics
+  (Section~\ref{sec:capab-mach-with})
+\item Define a novel calling convention enforcing control-flow correctness and
+  encapsulation of local state on the stack
+  (Section~\ref{sec:stack-and-return-pointer}). It relies solely on local
+  capabilities and does not require OS support (like a trusted stack or
+  call/return instructions). It also does not use separate per-component stacks
+  but a single, fairly standard, contiguous stack. It supports higher-order
+  cross-component calls (e.g., (potentially) cross-component function pointers)
+  and can be efficient assuming only one additional piece of processor support:
+  an efficient instruction for clearing a memory range.
+\item Prove a theorem that is the most general and powerful formulation of the
+  guarantees offered by a capability machine (a form of capability
+  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
+  guarantees offered for local capabilities. It is very general and not tied to
+  our calling convention or a specific way of using the system's capabilities.
+\item Introduce two novel technical ideas in the unary, step-indexed Kripke
+  logical relation used to formulate the above theorem. First, we model the
+  naturally continuation-passing-style of assembly code (with return pointers as
+  continuations) using a \emph{single} orthogonal closure, rather than the
+  existing biorthogonal closure. Secondly, we express the special nature of
+  local capabilities using a variant of Dreyer et. al.'s public and private
+  future worlds~\citep{Dreyer:jfp12}, used to express well-bracketedness
+  guarantees in a lambda calculus. The logical relation and capability safety
+  theorem is presented in Section~\ref{sec:logical-relation}.
+\item Demonstrate our results by applying them to challenging examples,
+  specifically constructed to demonstrate local state encapsulation and
+  control-flow correctness guarantees in the presence of cross-component
+  function pointers (Section~\ref{sec:examples}). The examples demonstrate both
+  the power of our formulation of capability safety and our calling convention.
 \end{itemize}
@@
 \label{sec:capab-mach-with}
-In this paper, we work with a formal capability machine that has all the
+In this paper, we work with a formal capability machine with all the
 characteristics of real capability machines, as well as local capabilities much
-like CHERI's, but is otherwise kept as simple as possible. It is inspired by
+like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
 both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
-CHERI~\cite{Watson2015Cheri}. To avoid getting bogged down in uninteresting
-detail, we assume an infinite address space and unbounded integers.
+CHERI~\cite{Watson2015Cheri}. To avoid uninteresting detail, we assume an
+infinite address space and unbounded integers.
 
@@
 One of the contributions in this paper is a demonstration that local
-capabilities on a capability machine can be used in a stack and return pointer
-management scheme that enforces control-flow correctness in a way that is provably
-watertight, potentially efficient, does not rely on a trusted central stack
-manager, and with support for exposing higher-order interfaces to an adversary.
-In this section, we explain this discipline. We start by explaining the
-high-level approach, the different issues that need to be taken into account,
-starting with first-order interfaces and moving onto a higher-order setting
-next. After that, we define a number of reusable macro-instructions that
-can be used to conveniently apply the proposed discipline in subsequent
-examples.
+capabilities on a capability machine can be used in a calling convention that
+enforces control-flow correctness in a way that is provably watertight,
+potentially efficient, does not rely on a trusted central stack manager, and
+with support for exposing higher-order interfaces to an adversary. In this
+section, we explain this convention. We start by explaining the high-level
+approach, the different issues that need to be taken into account, starting with
+first-order interfaces and moving onto a higher-order setting next. After that,
+we define a number of reusable macro-instructions that can be used to
+conveniently apply the proposed convention in subsequent examples.
 
@@
 previous invocations and then use them when they're no longer supposed to. To
-prevent this, before returning, we need to clear (i.e.\ overwrite with zeros) the
-entire part of the stack that the adversary has had access to, not just the
+prevent this, before returning, we need to clear (i.e.\ overwrite with zeros)
+the entire part of the stack that the adversary has had access to, not just the
 parts that we have used ourselves. Essentially, this is needed for the same
@@
 and back to the adversary. Clearing the entire unused part of the stack is the
-most problematic aspect of our discipline in terms of efficiency, since it may
-consist of a large part of memory. However, we believe there are ways to
-optimize it that we discuss in Section~\ref{sec:discussion}.
+most problematic aspect of our calling convention in terms of efficiency, since
+it may consist of a large part of memory (but see Section~\ref{sec:discussion}
+for how this might be mitigated).
 
@@
 placing and executing instructions on the stack, i.e.\ it will not just contain
-code pointers and data. This means that our discipline should be combined with
-protection against stack smashing attacks (i.e.\ buffer overflows on the stack
-overwriting activation records' contents), but conveniently, the fine-grained
-memory protection on capability machines is well-suited to easily provide such
-protection.
+code pointers and data. This means that our calling convention should be
+combined with protection against stack smashing attacks (i.e.\ buffer overflows
+on the stack overwriting activation records' contents), but conveniently, the
+fine-grained memory protection on capability machines is well-suited to easily
+provide such protection.
 
@@
 
-\textbf{Stack management discipline}
-The precautions the needed to safely interact with an adversary have now been introduced and motivated. The stack discipline comprises the precautions. Summarized, they were:
+\textbf{Calling convention} With the security measures introduced and motivated,
+here is a summary of our proposed calling convention:
 \begin{description}[font=\normalfont\itshape]
-\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
-\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used (any part of the stack that has been passed to an adversary must also be erased).
-\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in a register, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
-\item[Before invoking an adversary callback] make sure it is a global capability.
-\item[When called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
+\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
+\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
+\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
+\item[Before invoking an adversary callback] make sure the callback capability is global.
+\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
 \end{description}
 \textbf{Reusable macro instructions} 
-We define a number of reusable macros. The macros capture the stack
-management discipline and other conveniences. The macros are:
+We define a number of reusable macros capturing the calling convention and other conveniences. The macros are:
 \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
@@
 the adversary and prepares the received stack pointer by checking that
-it has permission $\rwlx$ and making it conform to our stack
-discipline. \texttt{\footnotesize{scall $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} Jumps
+it has permission $\rwlx$ and making it conform to our calling convention. \texttt{\footnotesize{scall $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} Jumps
 to the capability in register $r$ in the manner described above. That
@@
 to properly check the stack pointer received by our closure from the adversary
-(it must have the correct permission and fit with our stack discipline) and (2)
+(it must have the correct permission and fit with our calling convention) and (2)
 require that the adversary callback is global.
@@
 stack pointer or trick the trusted code into preserving it across an invocation,
-the assertion can likely be made to fail too. However, our stack and return
-pointer discipline guarantees that there is no way any of this can happen, as we
-prove in the following lemma.
+the assertion can likely be made to fail too. However, our calling convention
+guarantees that there is no way any of this can happen, as we prove in the
+following lemma.
 
@@
 
-\subsection{Stack and return pointer discipline}
+\subsection{Calling convention}
 \textbf{Formulating control flow correctness} The attentive reader may notice
-that, while we claim that our stack and return pointer discipline enforces
-control-flow correctness, we do not prove a general theorem to show that this is
-the case. The reason for this is that it is not clear what such a theorem should
-look like. Formulations in terms of a control-flow graph, like the one by
+that, while we claim that our calling convention enforces control-flow
+correctness, we do not prove a general theorem to show that this is the case.
+The reason for this is that it is not clear what such a theorem should look
+like. Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
-like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on. In fact, our
-examples show that our logical relation imply a form of control-flow correctness
-that is stronger than such formulations, although this is not made very
-explicit. As future work, we are considering to look at a more explicit and
-useful way to formalise well-bracketed control flow. The idea would be to define
-a variant of our capability machine with built-in call and return instructions
-and well-bracketed control flow built-in to the operational semantics, and then
-prove that compiling such code to our machine using our stack and return pointer
-discipline is fully abstract~\cite{abadi_protection_1998}, which roughly means
-that the lower-level machine does not allow interactions with the compiled code
-that do not correspond to an interaction with the original code in the
+like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
+In fact, our examples show that our logical relation imply a form of
+control-flow correctness that is stronger than such formulations, although this
+is not made very explicit. As future work, we are considering to look at a more
+explicit and useful way to formalise well-bracketed control flow. The idea would
+be to define a variant of our capability machine with built-in call and return
+instructions and well-bracketed control flow built-in to the operational
+semantics, and then prove that compiling such code to our machine using our
+calling convention is fully abstract~\cite{abadi_protection_1998}, which roughly
+means that the lower-level machine does not allow interactions with the compiled
+code that do not correspond to an interaction with the original code in the
 high-level machine.
@@
 \textbf{Performance and the requirement for stack clearing} It is important to
-consider the performance of our proposed stack and return management discipline.
+consider the performance of our proposed calling convention.
 Generally, the additional security measures described in
@@
 
-\textbf{Modularity} An important aspect of our stack and return pointer
-discipline is that it is modular. By that, we mean that the requirements we have
-on callbacks, return pointers etc. provided to trusted code by an adversary are
-also satisfied by callbacks, return pointers etc. that we construct ourselves.
-For example, our return pointers are necessarily local capabilities because they
-must point to memory where we can store the old stack pointer, but the
-adversary's return pointers are also allowed to be local. Adversary callbacks on
-the other hand are required to be global but the callbacks we construct
-ourselves are also allocated on the heap and global. In other words, we do not
-assume that our code has some sort of special status and it is possible for the
-adversary to protect themselves against us in the same way as we protect against
-them.
+\textbf{Modularity} An important aspect of our calling convention is that it is
+modular. By that, we mean that the requirements we have on callbacks, return
+pointers etc. provided to trusted code by an adversary are also satisfied by
+callbacks, return pointers etc. that we construct ourselves. For example, our
+return pointers are necessarily local capabilities because they must point to
+memory where we can store the old stack pointer, but the adversary's return
+pointers are also allowed to be local. Adversary callbacks on the other hand are
+required to be global but the callbacks we construct ourselves are also
+allocated on the heap and global. In other words, we do not assume that our code
+has some sort of special status and it is possible for the adversary to protect
+themselves against us in the same way as we protect against them.
 
@@
 \todo[inline]{Insert the correct URL for the technical report in the bibtex file!}
+
 \begin{comment}
@@
 \end{comment}
+
 %\bibliographystyle{IEEEtran}

### 2017-02-15 csf-2017/paper.tex
merge and first paragraph of discussion

@@
 \label{sec:discussion}
-
 \subsection{Calling convention}
-\textbf{Formulating control flow correctness} The attentive reader may notice
-that, while we claim that our calling convention enforces control-flow
-correctness, we do not prove a general theorem to show that this is the case.
-The reason for this is that it is not clear what such a theorem should look
-like. Formulations in terms of a control-flow graph, like the one by
+\textbf{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
+correctness, we do not prove a general theorem that shows this, it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
@@
 control-flow correctness that is stronger than such formulations, although this
-is not made very explicit. As future work, we are considering to look at a more
+is not made very explicit. As future work, we consider looking at a more
 explicit and useful way to formalise well-bracketed control flow. The idea would

### 2017-02-15 csf-2017/paper.tex
Shortening, small typesetting changes.

--- 2017-02-15 csf-2017/paper.tex
@@
     \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\perm,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ \footnotesize{ otherwise }$\var{newPc} = \memreg(r)$\\
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline
+    \footnotesize{otherwise }$\var{newPc} = \memreg(r)$\\
     \hline
@@
 
-We define the syntax of our capability machine in
-\figurename~\ref{fig:syntax}. We assume an infinite set of addresses
-$\Addrs$ and define machine words as either integers or capabilities
-of the form $((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability
-represents the authority to execute permissions $\perm$ on the memory
-array $[\var{base},\var{end}]$, together with a current address $\addr$ and
-a tag $\gl$ which indicates whether the capability is global or
-local. The capability machine has no notion of pointers by itself, so
-we refer to capabilities as pointers and capabilities
-interchangeably. The available permissions and their ordering are
-depicted in \figurename~\ref{fig:perm-hier}: the permissions include
-null permission ($\noperm$), readonly ($\readonly$), read/write
-($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$)
-permissions. Additionally, there are three special read/write-local
-($\readwritel$), read/write-local/execute ($\rwlx$) and enter
-($\entry$) permissions which we will explain below.
-
-We assume a set of register names $\RegName$. We define register files $\reg$ as
-functions mapping register names to words and memories as functions that map
-addresses to words. The state of the entire machine is represented as a
-configuration that is either a running $\Phi \in \ExecConfs$ containing a memory
-and a register file, or a failed or halted state, where the latter keeps hold
-of the final state of memory. 
+We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
+We assume an infinite set of addresses $\Addrs$ and define machine words as
+either integers or capabilities of the form
+$((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability represents the
+authority to execute permissions $\perm$ on the memory range
+$[\var{base},\var{end}]$, together with a current address $\addr$ and a tag
+$\gl$ indicating whether the capability is global or local. There is no notion
+of pointers other than capabilities, so we will use the terms interchangeably.
+The available permissions and their ordering are depicted in
+\figurename~\ref{fig:perm-hier}: the permissions include null permission
+($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
+($\exec$) and read/write/execute ($\rwx$) permissions. Additionally, there are
+three special permissions: read/write-local ($\readwritel$), read/write-local/execute
+($\rwlx$) and enter ($\entry$), which we will explain below.
+
+We assume a finite set of register names $\RegName$. We define register files
+$\reg$ and memories $\ms$ as functions mapping register names resp. addresses to
+words. The state of the entire machine is represented as a configuration that is
+either a running $\Phi \in \ExecConfs$ containing a memory and a register file,
+or a failed or halted state, where the latter keeps hold of the final state of
+memory.
 % We also define memory segments $\ms$:
@@
 
-The instruction set of our machine is rather basic. Instructions $i$ include
+The machine's instruction set is rather basic. Instructions $i$ include
 relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
@@
 load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and an arithmetic addition operators ($\plus{}{}{}$ and $\minus{}{}{}$). There are
-three instructions for modifying capabilities: $\lea{}{}$ (modifies the current
-address), $\restricttwo{}{}$ (modifies the permission and local/global flag) and
+and $\store{}{}$) and arithmetic addition operators ($\plus{}{}{}$ and
+$\minus{}{}{}$, operating only on numbers). There are three instructions for
+modifying capabilities: $\lea{}{}$ (modifies the current address),
+$\restricttwo{}{}$ (modifies the permission and local/global flag) and
 $\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
@@
 authority than the original (e.g. restrict will only weaken a permission).
-Finally, there is an instruction $\isptr{}{}$ for testing whether a word is a
-capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$,
-$\gete{}{}$ and $\geta{}{}$ for obtaining respectively the permissions,
-local/global flag, base, end and current address of a capability.
+Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a
+number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and
+$\geta{}{}$ provide access to a capability's permissions, local/global flag, base,
+end and current address, respectively.
 
@@
 \end{comment}
-\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics containing a number of representative
-instructions. Essentially, a configuration $\Phi$ will either decode and execute
-the instruction at $\memreg{\pcreg}$ if it is executable and the current address
-is in the valid range or fail otherwise. The table in the figure shows for a
-given instruction $i$ the result of executing it in configuration $\Phi$. The
-$\fail$ and $\halt$ instructions simply fail and halt. The $\move{}{}$
-instruction simply modifies the register file as requested and updates the
-$\pcreg$ to the next instruction using the meta-function $\stdUpdatePc{}$.
+\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics for a
+number of representative instructions. Essentially, a configuration $\Phi$ will
+either decode and execute the instruction at $\memreg{\pcreg}$ if it is
+executable and the current address is in the valid range or fail otherwise. The
+table in the figure shows for a given instruction $i$ the result of executing it
+in configuration $\Phi$. The $\fail$ and $\halt$ instructions obviously fail and
+halt respectively. The $\move{}{}$ instruction simply modifies the register file
+as requested and updates the $\pcreg$ to the next instruction using the
+meta-function $\stdUpdatePc{}$.
 
@@
 suitable for representing a kind of closures: an opaque package containing a
-piece of code together with local encapsulated state. The idea is that one can
-build an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where the
-range $[\start,\addr-1]$ is used for storing the closure's local state (data or
-capabilities) and the range $[\addr,\addrend]$ can be used for storing
-instructions. The instructions and data are opaque to an adversary holding $c$
-but when $c$ is jumped to, the data becomes available to the closure's code through
-the updated version of $c$ that is now in the $\pcreg$\footnote{Note that the
-  capability in the $\pcreg$-register can be used to load from.}.
+piece of code together with local encapsulated state. Such a package can be
+built as an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where
+the range $[\start,\addr-1]$ contains local state (data or capabilities) and
+$[\addr,\addrend]$ contains instructions. The package is opaque to an adversary
+holding $c$ but when $c$ is jumped to, the instructions can start executing and
+have access to the local data through the updated version of $c$ that is then in
+$\pcreg$\footnote{Note that the capability in the $\pcreg$-register can be used
+  to load from.}.
 % \lau{As we

### 2017-02-15 csf-2017/paper.tex
Discussion first pass

--- 2017-02-15 csf-2017/paper.tex
@@
 capabilities left there by them, including local capabilities since the stack is
-store-local. The adversary could take advantage of this by using the unused part
+write-local. The adversary could take advantage of this by using the unused part
 of the stack to store either stack pointers or return pointers from such
@@
 In large parts, capability safety is defined by when a capability
-can be safely given to an untrusted piece of code. In the following paragraphs,
+can be safely given to an adveserial piece of code. In the following paragraphs,
 we will go through the definitions of the logical relation and provide an
@@
 The Fundamental Theorem allows us to prove capability safety for
-untrusted programs: if the authority granted by
+adveserial programs: if the authority granted by
 the capabilty for the untrustred program is reasonable, then it is safe to
@@
 to store its local state. It also uses the stack-based calling convention
-\texttt{\footnotesize{scall}} to call the untrusted code. On the other hand, \texttt{\footnotesize{f2}} uses
+\texttt{\footnotesize{scall}} to call the adversary. On the other hand, \texttt{\footnotesize{f2}} uses
 malloc to allocate memory for its local state and uses an activation-record
@@
 
-\textbf{Performance and the requirement for stack clearing} It is important to
-consider the performance of our proposed calling convention.
-Generally, the additional security measures described in
-Section~\ref{sec:stack-and-return-pointer} obviously impose an overhead w.r.t.
-an approach that does not provide security. However, they generally consist of a
+\textbf{Performance and the requirement for stack clearing} 
+The additional security measures of the calling convention described in
+Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to
+a calling convention that does not provide security. However, the security measures generally consist of a
 few checks or register clearings on every boundary crossing between trusted code
 and adversary, which should impose an acceptable performance overhead. The only
-exception is the requirements that we have for stack clearing, in two
+exception is the requirements for stack clearing which we occur in two
 situations: when returning to the adversary and when invoking an adversary
 callback. As we have explained, it is not sufficient to clear just the part of
-the stack that we have used ourselves, but we need to clear the entire stack,
-except the part that we are using ourselves. This means that on every crossing
-of the boundary between trusted and adversary code, a potentially large region
-of memory needs to be zeroed out entirely. We believe that this in fact is a
-common requirement for typical usage scenario's of local capabilities and
+the stack that we have used ourselves, but we need to clear all of the stack that we are not using ourselves. 
+In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be erased.
+We believe this is a common requirement for typical usage scenarios of local capabilities and
 capability machines like CHERI should consider to provide special support for
-this requirement, in the form of a highly-optimized instruction for zeroing out
-a potentially large block of memory. We have no experience in processor design,
+this requirement, in the form of a highly-optimized instruction for erasing
+a large block of memory. We have no experience in processor design,
 but we imagine this might be optimized as a vector instruction or even further,
@@
 
-\textbf{Modularity} An important aspect of our calling convention is that it is
-modular. By that, we mean that the requirements we have on callbacks, return
+\textbf{Modularity} It is important that our calling convention is modular.
+This means that the requirements we have on callbacks, return
 pointers etc. provided to trusted code by an adversary are also satisfied by
 callbacks, return pointers etc. that we construct ourselves. For example, our
-return pointers are necessarily local capabilities because they must point to
+return pointers are local capabilities because they must point to
 memory where we can store the old stack pointer, but the adversary's return
@@
 has some sort of special status and it is possible for the adversary to protect
-themselves against us in the same way as we protect against them.
+themselves against us the same way we protect against them.
 
@@
 world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
-the awkward example uses exactly the same state transition system than theirs),
+the awkward example uses exactly the same state transition system),
 but they turn up in an entirely different place in the logical relation: rather
 than using public future worlds for the special syntactic category of evaluation
-contexts, they are used in the value relation depending on the depending on the
+contexts, they are used in the value relation depending on the
 locality of the capability at hand. Additionally, our worlds are a bit more
-complex because, to allow local memory capabilities and store-local
+complex because, to allow local memory capabilities and write-local
 capabilities, they can contain (revokable) temporary regions that are only
@@
 immediate use without allowing it to be stored in a closure or reference for
-later use. We believe reasoning about such values will require very similar
-reasoning techniques than the ones we have used for local capabilities.
+later use. We believe reasoning about such values will require techniques similar 
+to the ones we have used for local capabilities.
 

### 2017-02-15 csf-2017/paper.tex
first pass related work

@@
 previous work in the area. The capability machine formalised in
-Section~\ref{sec:capab-mach-with} is a simple but representative model, modeled
+Section~\ref{sec:capab-mach-with} is a simple but representative model. It is modeled
 mainly after the M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter
@@
 enforce only partial correctness of control flow. This includes particularly the
-line of work on \emph{control flow integrity}~\citep{abadi_control-flow_2005}.
+line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
 This work uses a quite different attacker model than us: they assume an attacker
@@
 every indirect jump and using memory access control to prevent overwriting code,
-this work enforces what they call control flow integrity, formalised as the
-property that every jump will follow a legal path in the control flow graph.
+this work enforces what they call control-flow integrity, formalised as the
+property that every jump will follow a legal path in the control-flow graph.
 As discussed in Section~\ref{sec:discussion}, this correctness property does not
@@
 capability safety in a high-level JavaScript-like lambda calculus. Our
-fundamental theorem is similar to theirs and can be seen as a formulation of
+Fundamental Theorem is similar to theirs and can be seen as a formulation of
 capability safety of the capability machine. Because we are not interested in
@@
 result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
-correctness). He also proves soundness of control flow integrity (see above)
+correctness). He also proves soundness of control-flow integrity (see above)
 adapted to the capability machine, a result that seems orthogonal to the fact
@@
 %\section{Conclusion}
-\label{sec:conclusion}
+%\label{sec:conclusion}
 \todo[inline]{Insert the correct URL for the technical report in the bibtex file!}

### 2017-02-15 csf-2017/paper.tex
Discussion, cutting explanation of full abstraction and speculative part about memory erasure instruction.

@@
 semantics, and then prove that compiling such code to our machine using our
-calling convention is fully abstract~\cite{abadi_protection_1998}, which roughly
-means that the lower-level machine does not allow interactions with the compiled
-code that do not correspond to an interaction with the original code in the
-high-level machine.
+calling convention is fully abstract~\cite{abadi_protection_1998}.
 
@@
 this requirement, in the form of a highly-optimized instruction for erasing
-a large block of memory. We have no experience in processor design,
-but we imagine this might be optimized as a vector instruction or even further,
-by having the processor keep track of a range of memory for which loads should
-pretend that it has already been cleared, by returning zero irrespective of the
-actual contents of the memory.
+a large block of memory.
 

### 2017-02-15 csf-2017/paper.tex
Shorten.

--- 2017-02-15 csf-2017/paper.tex
@@
 
-The final semantics we discuss is that of $\store{}{}$. As expected, it updates
-the memory to the requested value if the capability has write authority for the
-requested location. However, the instruction is complicated by the presence of
-\emph{local capabilities}, modeled after the ones in
-the CHERI processor~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
-they can only be kept in registers, i.e.\ they cannot be stored to memory. This
-means that local capabilities can be \emph{temporarily} given to an adversary,
-for the duration of an invocation: if we take care to erase the capability from
-the register file after control is passed back to us, we can be sure that they
-will not have been able to store the capability. However, there is one exception
-to the rule above: local capabilities can be stored to a part of memory for
-which we have a capability with write-local authority (i.e.\ permission $\rwl$ or
-$\rwlx$). This exception is intended to accomodate regions of memory that behave
-like the C stack: stashing away the contents of registers on the stack can be
-allowed, even if this includes local capabilities. As long as all capabilities
-with write-local authority are themselves local and the stack is cleared after
-control is passed back by the adversary, we will see that this does not break
-the intended behavior of local capabilities.
+Finally, the $\store{}{}$ instruction updates the memory to the requested value
+if the capability has write authority for the requested location. However, the
+instruction is complicated by the presence of \emph{local capabilities}, modeled
+after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
+capabilities are special in that they can only be kept in registers, i.e.\ they
+cannot be stored to memory. This means that local capabilities can be
+\emph{temporarily} given to an adversary, for the duration of an invocation: if
+we take care to erase the capability from the register file after control is
+passed back to us, they will not have been able to store the capability.
+However, there is one exception to the rule above: local capabilities can be
+stored to memory for which we have a capability with write-local authority
+(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accomodate a stack,
+where register contents can be stored, including local capabilities. As long as all
+capabilities with write-local authority are themselves local and the stack is
+cleared after control is passed back by the adversary, we will see that this
+does not break the intended behavior of local capabilities.
 
 We point out that our local capabilities capture only a part of the semantics of
-local capabilities in CHERI. Specifically, on top of the above restrictions,
-CHERI's default implementation of CCall additionally forbids local capabilities
-from being passed accross module boundaries, something which we rely on
-crucially in this paper, for passing around local return pointers and stack
-capabilities. However, CHERI's CCall is not implemented in hardware, but in
-software, precisely to allow experimenting with alternative models like ours.
+local capabilities in CHERI. Specifically, in addition to the above, CHERI's
+default implementation of the CCall exception handler forbids local capabilities
+from being passed accross module boundaries. Such a restriction fundamentally
+breaks our calling convention, since we pass around local return pointers and
+stack capabilities. However, CHERI's CCall is not implemented in hardware, but
+in software, precisely to allow experimenting with alternative models like ours.
 
@@
 
-In order to have a reasonably realistic system, we have created a
-simple model of linking, so programs can have access to other programs
-through a linking table. We also assume malloc to be part of the
-trusted computing base satisfying a certain specification. Malloc and
-the model linking is described to the degree necessary, in the next
-section, but we refer to the technical appendix~\cite{technical_appendix} for further
-details.
+In order to have a reasonably realistic system, we use a simple model of
+linking, so programs can have access to other programs through a linking table.
+We also assume malloc to be part of the trusted computing base satisfying a
+certain specification. Malloc and linking tables are described further in the
+next section, but we refer to the technical appendix~\cite{technical_appendix}
+for full details.
 
@@
 \label{sec:stack-and-return-pointer}
-%\lau{Do we want to use the term ``pointer'' at any point? I have also used it, but I guess it might create some confusing to whether we have traditional pointer as well as capabilities?}
 One of the contributions in this paper is a demonstration that local
-capabilities on a capability machine can be used in a calling convention that
-enforces control-flow correctness in a way that is provably watertight,
-potentially efficient, does not rely on a trusted central stack manager, and
-with support for exposing higher-order interfaces to an adversary. In this
-section, we explain this convention. We start by explaining the high-level
-approach, the different issues that need to be taken into account, starting with
-first-order interfaces and moving onto a higher-order setting next. After that,
-we define a number of reusable macro-instructions that can be used to
-conveniently apply the proposed convention in subsequent examples.
-
-The basic idea of our approach is quite simple: we stick to a single, rather
-standard, C stack and register-passed return pointers, much like a standard C
-ABI. However, to prevent various ways of
-misusing this basic scheme, we put local capabilities to work and take a number
-of not-always-obvious safety measures. In the next paragraphs, we will explain
-the issues to be considered in all the relevant situations: when (1) starting
-our program, (2) Returning to the adversary, (3) invoking the adversary, (4)
-returning from the adversary, (5) invoking an adversary callback and (6) having
-a callback invoked by the adversary.
-
-% \dominique{Add a summary table somewhere?}
-% \lau{It seems like we want have the space for this}.
-
-\textbf{Program startup}
-When our code starts executing, we assume that the language runtime initializes
-the memory as follows: a contiguous array of memory is reserved for the stack,
-for which we receive a stack pointer in a register, $r_\stk$, that we reserve for this
-purpose. We stress that the stack is not built-in, but merely an
-abstraction we put on this piece of the memorey. This stack pointer must be local and it must carry $\rwlx$ permission. It is crucial for
-our approach that the stack is the only part of memory for which there will ever
-be an obtainable capabilities with $\rwlx$ or $\rwl$ capabilities. We expect the language
-runtime to enforce this by not handing out such capabilities for other parts of
-memory in any situation (i.e.\ not through malloc, not during loading or linking
-etc.). Additionally, in our examples, we typically also assume some static parts
-of memory where instructions are stored or static data. Another part of memory
-(called the heap) is initially governed by malloc and at program startup, no
-other code has capabilities for this memory. Malloc will hand out capabilities
-for allocated regions as requested, but it will never hand out capabilities with
-write-local capability (i.e.\ no $\rwlx$ or $\rwl$ permissions). For simplicity,
-we assume that memory allocated through malloc cannot be freed.
-
-\textbf{Returning to the adversary} Perhaps the simplest scenario is the case of
-returning to the adversary after they invoked our code. In this case, we will
-have received a return pointer from them, and we just need to jump to it as
-usual. The only thing we need to take care of is to properly clear the registers
-before we do (erasing any data we do not want to pass as (part of) the return
-value) and similarly. However, it is important to carefully consider what needs
-to happen with the stack. Since we may have used the stack for various purposes
-(register spilling, storing local state when invoking other functions etc.), we
-clearly need to erase that data before returning to the adversary, to prevent
-the adversary access to it.
+capabilities on a capability machine support a calling convention that enforces
+control-flow correctness in a way that is provably watertight, potentially
+efficient, does not rely on a trusted central stack manager, and supports
+higher-order interfaces to an adversary. In this section, we explain this
+convention's high-level approach, the security measures to be taken in a number
+of situations (motivating each separately with a summary table at the end).
+After that, we define a number of reusable macro-instructions that can be used
+to conveniently apply the proposed convention in subsequent examples.
+
+The basic idea of our approach is simple: we stick to a single, rather standard,
+C stack and register-passed return pointers, much like a standard C calling
+convention. However, to prevent various ways of misusing this basic scheme, we
+put local capabilities to work and take a number of not-always-obvious safety
+measures. In the next paragraphs, we will explain the issues to be considered in
+all the relevant situations: when (1) starting our program, (2) returning to the
+adversary, (3) invoking the adversary, (4) returning from the adversary, (5)
+invoking an adversary callback and (6) having a callback invoked by the
+adversary.
+
+\textbf{Program startup} We assume that the language runtime initializes the
+memory as follows: a contiguous array of memory is reserved for the stack, for
+which we receive a stack pointer in a special register $r_\stk$. We stress that
+the stack is not built-in, but merely an abstraction we put on this piece of the
+memory. The stack pointer is local and has $\rwlx$ permission. Crucially, the
+stack is the only part of memory for which the runtime (including malloc,
+loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
+Additionally, our examples typically also assume some memory to store
+instructions or static data. Another part of memory (called the heap) is
+initially governed by malloc and at program startup, no other code has
+capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
+regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
+assume that memory allocated through malloc cannot be freed.
+
+\textbf{Returning to the adversary} Perhaps the simplest situation is returning
+to the adversary after they invoked our code. In this case, we have received a
+return pointer from them, and we just need to jump to it as usual. An obvious
+security measure to take care of is properly clearing the non-return-value
+registers before we jump (since they may contain data or capabilities that the
+adversary should not get access to). Additionally, we may have used the stack
+for various purposes (register spilling, storing local state when invoking other
+functions etc.), so we also need to erase that data before returning to the
+adversary.
 
 However, if we are returning from a function that has itself invoked adversary
-code, then even the unused part of the stack may still contain data and
-capabilities left there by them, including local capabilities since the stack is
-write-local. The adversary could take advantage of this by using the unused part
-of the stack to store either stack pointers or return pointers from such
-previous invocations and then use them when they're no longer supposed to. To
-prevent this, before returning, we need to clear (i.e.\ overwrite with zeros)
-the entire part of the stack that the adversary has had access to, not just the
-parts that we have used ourselves. Essentially, this is needed for the same
-reason that we clear the register file: because the adversary should not be able
-to keep hold of local capabilities when control is passed to the trusted code
-and back to the adversary. Clearing the entire unused part of the stack is the
-most problematic aspect of our calling convention in terms of efficiency, since
-it may consist of a large part of memory (but see Section~\ref{sec:discussion}
-for how this might be mitigated).
-
-\textbf{Invoking the adversary} Slightly more complex is the case where we
-invoke the adversary. As above, we clear all the registers, except those that
-contain arguments the adversary is intended to have access to. We leave a copy
-of the stack pointer in $r_\stk$, but only after we have used
-the $\subseg{}{}{}$ to shrink its authority to just the part of the stack that
-we are not using. In one of the registers, we also provide a return pointer,
-which we will say more about below.
-
-Already important is the fact that the return pointer must be provided to the
-adversary as a local capability. If not, the adversary would be able to store
-away the return pointer in a global data structure, and they would be able to jump
-to it in circumstances where this should not be possible: for example, they may
-store the return pointer, legally jump to it a first time, wait to be invoked
-again and then jump to the old return pointer a second time, rather than the new
-return pointer received for the new invocation. Similarly, they may store the
-return pointer, invoke a function in our code, wait for us to invoke him again
-and then jump to the stored, old return pointer rather than the new one they
-received upon the second invocation.
+code, then clearing the used part of the stack is not enough. The \emph{unused}
+part of the stack may also contain data and capabilities, left there by the
+adversary, including local capabilities since the stack is store-local. The
+adversary could take advantage of this by using the unused part of the stack to
+store either stack pointers or return pointers from such previous invocations
+and then use them when they're no longer supposed to. To prevent this, we need
+to clear (i.e.\ overwrite with zeros) the entire part of the stack that the
+adversary has had access to, not just the parts that we have used ourselves.
+This requirement may seem strong, but it is essentially there for the same
+reason that we clear the register file: the adversary should not be able to keep
+hold of local capabilities when control is passed to trusted code and back to
+the adversary. Since we may be talking about a large part of memory, this is the
+most problematic aspect of our calling convention for performance, but see
+Section~\ref{sec:discussion} for how this might be mitigated.
+
+\textbf{Invoking the adversary} A slightly more complex case is invoking the
+adversary. As above, we clear all the non-argument registers, as well as the
+part of the stack that we are not using (because, as above, it may contain local
+capabilities from previously executed code that the adversary could exploit in
+the same way as above). We leave a copy of the stack pointer in $r_\stk$, but
+only after we have used the $\subseg{}{}{}$ to shrink its authority to the part
+that we are not using ourselves.
+
+In one of the registers, we also provide a return pointer, which must be a local
+capability. If it were global, the adversary would be able to store away the
+return pointer in a global data structure, and jump to it later, in
+circumstances where this should not be possible. For example, they could store
+the return pointer, legally jump to it a first time, wait to be invoked again
+and then jump to the old return pointer a second time, instead of the new return
+pointer received for the second invocation. Similarly, they could store the
+return pointer, invoke a function in our code, wait for us to invoke them again
+and then jump to the old return pointer rather than the new one, received for the
+second invocation.
 % \lau{If we need space, then maybe one of the examples is sufficient.}
@@
 pointer.
-
-A final required safety measure is that before invocation of the adversary, we
-need to clear the part of the stack that we are not using. Otherwise, similarly
-to what we explained above for returning to the adversary, that part of the
-stack may still contain local capabilities from previously executed code and the
-adversary could exploit this in much the same way as above.
 
@@
 In large parts, capability safety is defined by when a capability
-can be safely given to an adveserial piece of code. In the following paragraphs,
+can be safely given to an untrusted piece of code. In the following paragraphs,
 we will go through the definitions of the logical relation and provide an
@@
 The Fundamental Theorem allows us to prove capability safety for
-adveserial programs: if the authority granted by
+untrusted programs: if the authority granted by
 the capabilty for the untrustred program is reasonable, then it is safe to
@@
 to store its local state. It also uses the stack-based calling convention
-\texttt{\footnotesize{scall}} to call the adversary. On the other hand, \texttt{\footnotesize{f2}} uses
+\texttt{\footnotesize{scall}} to call the untrusted code. On the other hand, \texttt{\footnotesize{f2}} uses
 malloc to allocate memory for its local state and uses an activation-record
@@
 \label{sec:discussion}
+
 \subsection{Calling convention}
-\textbf{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
-correctness, we do not prove a general theorem that shows this, it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
+\textbf{Formulating control flow correctness} The attentive reader may notice
+that, while we claim that our calling convention enforces control-flow
+correctness, we do not prove a general theorem to show that this is the case.
+The reason for this is that it is not clear what such a theorem should look
+like. Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
@@
 control-flow correctness that is stronger than such formulations, although this
-is not made very explicit. As future work, we consider looking at a more
+is not made very explicit. As future work, we are considering to look at a more
 explicit and useful way to formalise well-bracketed control flow. The idea would
@@
 
-\textbf{Performance and the requirement for stack clearing} 
-The additional security measures of the calling convention described in
-Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to
-a calling convention that does not provide security. However, the security measures generally consist of a
+\textbf{Performance and the requirement for stack clearing} It is important to
+consider the performance of our proposed calling convention.
+Generally, the additional security measures described in
+Section~\ref{sec:stack-and-return-pointer} obviously impose an overhead w.r.t.
+an approach that does not provide security. However, they generally consist of a
 few checks or register clearings on every boundary crossing between trusted code
 and adversary, which should impose an acceptable performance overhead. The only
-exception is the requirements for stack clearing which we occur in two
+exception is the requirements that we have for stack clearing, in two
 situations: when returning to the adversary and when invoking an adversary
 callback. As we have explained, it is not sufficient to clear just the part of
-the stack that we have used ourselves, but we need to clear all of the stack that we are not using ourselves. 
-In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be erased.
-We believe this is a common requirement for typical usage scenarios of local capabilities and
+the stack that we have used ourselves, but we need to clear the entire stack,
+except the part that we are using ourselves. This means that on every crossing
+of the boundary between trusted and adversary code, a potentially large region
+of memory needs to be zeroed out entirely. We believe that this in fact is a
+common requirement for typical usage scenario's of local capabilities and
 capability machines like CHERI should consider to provide special support for
-this requirement, in the form of a highly-optimized instruction for erasing
-a large block of memory. We have no experience in processor design,
+this requirement, in the form of a highly-optimized instruction for zeroing out
+a potentially large block of memory. We have no experience in processor design,
 but we imagine this might be optimized as a vector instruction or even further,
@@
 
-\textbf{Modularity} It is important that our calling convention is modular.
-This means that the requirements we have on callbacks, return
+\textbf{Modularity} An important aspect of our calling convention is that it is
+modular. By that, we mean that the requirements we have on callbacks, return
 pointers etc. provided to trusted code by an adversary are also satisfied by
 callbacks, return pointers etc. that we construct ourselves. For example, our
-return pointers are local capabilities because they must point to
+return pointers are necessarily local capabilities because they must point to
 memory where we can store the old stack pointer, but the adversary's return
@@
 has some sort of special status and it is possible for the adversary to protect
-themselves against us the same way we protect against them.
+themselves against us in the same way as we protect against them.
 
@@
 world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
-the awkward example uses exactly the same state transition system),
+the awkward example uses exactly the same state transition system than theirs),
 but they turn up in an entirely different place in the logical relation: rather
 than using public future worlds for the special syntactic category of evaluation
-contexts, they are used in the value relation depending on the
+contexts, they are used in the value relation depending on the depending on the
 locality of the capability at hand. Additionally, our worlds are a bit more
-complex because, to allow local memory capabilities and write-local
+complex because, to allow local memory capabilities and store-local
 capabilities, they can contain (revokable) temporary regions that are only
@@
 immediate use without allowing it to be stored in a closure or reference for
-later use. We believe reasoning about such values will require techniques similar 
-to the ones we have used for local capabilities.
+later use. We believe reasoning about such values will require very similar
+reasoning techniques than the ones we have used for local capabilities.
 

### 2017-02-15 csf-2017/paper.tex

@@
 semantics, and then prove that compiling such code to our machine using our
-calling convention is fully abstract~\cite{abadi_protection_1998}, which roughly
-means that the lower-level machine does not allow interactions with the compiled
-code that do not correspond to an interaction with the original code in the
-high-level machine.
-
-\textbf{Performance and the requirement for stack clearing} It is important to
-consider the performance of our proposed calling convention.
-Generally, the additional security measures described in
-Section~\ref{sec:stack-and-return-pointer} obviously impose an overhead w.r.t.
-an approach that does not provide security. However, they generally consist of a
+calling convention is fully abstract~\cite{abadi_protection_1998}.
+
+\textbf{Performance and the requirement for stack clearing} 
+The additional security measures of the calling convention described in
+Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to
+a calling convention that does not provide security. However, the security measures generally consist of a
 few checks or register clearings on every boundary crossing between trusted code
 and adversary, which should impose an acceptable performance overhead. The only
-exception is the requirements that we have for stack clearing, in two
+exception is the requirements for stack clearing which we occur in two
 situations: when returning to the adversary and when invoking an adversary
 callback. As we have explained, it is not sufficient to clear just the part of
-the stack that we have used ourselves, but we need to clear the entire stack,
-except the part that we are using ourselves. This means that on every crossing
-of the boundary between trusted and adversary code, a potentially large region
-of memory needs to be zeroed out entirely. We believe that this in fact is a
-common requirement for typical usage scenario's of local capabilities and
+the stack that we have used ourselves, but we need to clear all of the stack that we are not using ourselves. 
+In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be erased.
+We believe this is a common requirement for typical usage scenarios of local capabilities and
 capability machines like CHERI should consider to provide special support for
-this requirement, in the form of a highly-optimized instruction for zeroing out
-a potentially large block of memory. We have no experience in processor design,
-but we imagine this might be optimized as a vector instruction or even further,
-by having the processor keep track of a range of memory for which loads should
-pretend that it has already been cleared, by returning zero irrespective of the
-actual contents of the memory.
-
-\textbf{Modularity} An important aspect of our calling convention is that it is
-modular. By that, we mean that the requirements we have on callbacks, return
+this requirement, in the form of a highly-optimized instruction for erasing
+a large block of memory.
+
+\textbf{Modularity} It is important that our calling convention is modular.
+This means that the requirements we have on callbacks, return
 pointers etc. provided to trusted code by an adversary are also satisfied by
 callbacks, return pointers etc. that we construct ourselves. For example, our
-return pointers are necessarily local capabilities because they must point to
+return pointers are local capabilities because they must point to
 memory where we can store the old stack pointer, but the adversary's return

### 2017-02-15 csf-2017/paper.tex
reformatted subsections of discussion section

--- 2017-02-15 csf-2017/paper.tex
@@
 \label{sec:discussion}
-\subsection{Calling convention}
-\textbf{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
+\noindent\textbf{Calling convention}\\
+\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
 correctness, we do not prove a general theorem that shows this, it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
@@
 
-\textbf{Performance and the requirement for stack clearing} 
+\emph{Performance and the requirement for stack clearing} 
 The additional security measures of the calling convention described in
@@
 
-\textbf{Modularity} It is important that our calling convention is modular.
+\emph{Modularity} It is important that our calling convention is modular.
 This means that the requirements we have on callbacks, return
@@
 themselves against us the same way we protect against them.
-
-\subsection{Logical relation}
-
-\textbf{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in
+\\\\
+\noindent\textbf{Logical relation}\\
+\emph{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in
 Figure~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an
@@
 
-\textbf{Public/private future worlds} A novel aspect of our logical relation is
+\emph{Public/private future worlds} A novel aspect of our logical relation is
 how we model the temporary nature of local capabilities using public/private
@@
 
-\textbf{Local capabilities in high-level languages} It is worth mentioning that
+\emph{Local capabilities in high-level languages} It is worth mentioning that
 local capabilities are quite similar to a feature proposed for the high-level

### 2017-02-15 csf-2017/paper.tex
store-local => write-local.

--- 2017-02-15 csf-2017/paper.tex
@@
 part of the stack may also contain data and capabilities, left there by the
-adversary, including local capabilities since the stack is store-local. The
+adversary, including local capabilities since the stack is write-local. The
 adversary could take advantage of this by using the unused part of the stack to

### 2017-02-15 csf-2017/paper.tex
Shorten + clarify.

@@
 part of the stack may also contain data and capabilities, left there by the
-adversary, including local capabilities since the stack is write-local. The
-adversary could take advantage of this by using the unused part of the stack to
-store either stack pointers or return pointers from such previous invocations
-and then use them when they're no longer supposed to. To prevent this, we need
-to clear (i.e.\ overwrite with zeros) the entire part of the stack that the
-adversary has had access to, not just the parts that we have used ourselves.
-This requirement may seem strong, but it is essentially there for the same
-reason that we clear the register file: the adversary should not be able to keep
-hold of local capabilities when control is passed to trusted code and back to
-the adversary. Since we may be talking about a large part of memory, this is the
-most problematic aspect of our calling convention for performance, but see
-Section~\ref{sec:discussion} for how this might be mitigated.
+adversary, including local capabilities since the stack is write-local. As we
+will see later, we rely on the fact that the adversary cannot keep hold of local
+capabilities when they pass control to the trusted code and receive control
+back. In this case, the adversary could use the unused part of the stack to
+store local pointers and load them from there after he gets control back. To
+prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
+the stack that the adversary has had access to, not just the parts that we have
+used ourselves. Since we may be talking about a large part of memory, this
+requirement is the most problematic aspect of our calling convention for
+performance, but see Section~\ref{sec:discussion} for how this might be
+mitigated.
 
@@
 capabilities from previously executed code that the adversary could exploit in
-the same way as above). We leave a copy of the stack pointer in $r_\stk$, but
-only after we have used the $\subseg{}{}{}$ to shrink its authority to the part
-that we are not using ourselves.
+the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
+we have used the $\subseg{}{}{}$ to shrink its authority to the part that we are
+not using ourselves.
 
@@
 %   is so important that we do not want to leave it out.}
-
-Because the return pointer is local, this is not possible: the adversary can
-only store it through write-local capabilities, which (because of our
-assumptions above) means: on the stack. However, this makes it crucial that the
-stack pointer is also local: if the adversary were able to store the stack
-pointer in a global data structure, they can store return pointers on the
-stack, and play similar mischief as mentioned above. Storing away stack pointers
-for use during future invocations would also be dangerous for other reasons:
-during such a future invocation, a larger part of the stack may be in use by the
-trusted code, which the adversary may still have access to through the old stack
-pointer.
+By making the return pointer local, we prevent such attacks: the adversary can
+only store local capabilities through write-local capabilities, which (because
+of our assumptions above) means: on the stack. Since the stack pointer itself is
+also local, it can also only be stored on the stack. Because we clear the part
+of the stack that the adversary has had access to before we pass control back,
+there is no way for them to recover either of these local capabilities.
+
+Note that storing stack pointers for use during future invocations would also be
+dangerous in itself, i.e. not just because it can be used to store return
+pointers. Imagine that the adversary stores his stack pointer, invokes trusted
+code that uses part of the stack to store private data and then invokes the
+adversary again with a stack pointer restricted to exclude the part containing
+the private data. If the adversary had a way of keeping hold of their old stack
+pointer, it could access the private data stored there by the trusted code, and
+break local state encapsulation.
 

### 2017-02-15 csf-2017/paper.tex
added comment, fixed typos

--- 2017-02-15 csf-2017/paper.tex
@@
 part of the stack may also contain data and capabilities, left there by the
-adversary, including local capabilities since the stack is write-local. The
+adversary, including local capabilities since the stack is store-local. The
 adversary could take advantage of this by using the unused part of the stack to
@@
 \emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
-correctness, we do not prove a general theorem that shows this, it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
+correctness, we do not prove a general theorem that shows this, because it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
@@
 as the code returns to a stable state (i.e. transitions to a public future world
-of the original) before returning (i.e. invoking the continuation). We
+of the original) before returning (i.e.
+invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation).
+Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
+We
 generalise this idea to reason about local capabilities: validity of local

### 2017-02-15 csf-2017/paper.tex
a couple more changes to discussion and related work

@@
 Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
-\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview of
-previous work in the area. The capability machine formalised in
-Section~\ref{sec:capab-mach-with} is a simple but representative model. It is modeled
-mainly after the M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter
-pointers resemble the M-Machine's) and
-CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the memory
-capabilities resemble CHERI's). The latter is a recent and relatively mature
-capability machine, which combines capabilities with a virtual memory approach,
-in the interest of backwards compatibility and gradual adoption. Our local
-capabilities are inspired by CHERI's, although as discussed, they can be passed
-accross module boundaries, contrary to what is enforced by CHERI's default CCall
-implementation.
-
-There are plenty of other papers that enforce well-bracketed control flow at a
+\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
+of previous work. The capability machine formalised in
+Section~\ref{sec:capab-mach-with} is a simple but representative
+model, modeled mainly after the
+M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter pointers
+resemble the M-Machine's) and
+CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
+memory and local capabilities resemble CHERI's). The latter is a
+recent and relatively mature capability machine, which combines
+capabilities with a virtual memory approach, in the interest of
+backwards compatibility and gradual adoption. As discussed, our local
+capabilities differ from CHERI's by being able to cross module boundaries, 
+contrary to what is enforced by CHERI's default CCall implementation.
+
+There are plenty of other papers\lau{If there are plenty, then I guess we should cite more than one?}
+ that enforce well-bracketed control flow at a
 low level, but most are restricted to preventing particular types of attacks and
@@
 property that every jump will follow a legal path in the control-flow graph.
-As discussed in Section~\ref{sec:discussion}, this correctness property does not
+As discussed in Section~\ref{sec:discussion}, this integrity property does not
 take into account temporal properties and seems hard to use for reasoning.
@@
 
-Our logical relation is a unary, step-indexed Kripke logical relation (SKLR)
+Our logical relation is a unary, step-indexed Kripke logical relation
 with recursive
@@
 containing state transition systems~\citep{Ahmed:popl09}, public/private future
-worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion),
+worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
 biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

### 2017-02-15 csf-2017/paper.tex
Shorten, clarify.

--- 2017-02-15 csf-2017/paper.tex
@@
 answer the last question first by considering what should happen when the
-adversary jumps to a return pointer and what capabilities we need for achieving
-that. In that case, the program counter should be restored to point after the
-instruction that jumped into the adversary, so the activation record should
-store this old program counter. Additionally, the stack pointer should also be
-restored to its original value. Since the adversary has a more restricted
-authority over the stack than the code making the call, we cannot hope to
-reconstruct the return pointer from the stack pointer owned by the adversary.
-Instead, the old stack pointer should also be stored as part of the activation
-record.
+adversary jumps to a return pointer. In that case, the program counter should be
+restored to the instruction after the jump to the adversary, so the activation
+record should store this old program counter. Additionally, the stack pointer
+should also be restored to its original value. Since the adversary has a more
+restricted authority over the stack than the code making the call, we cannot
+hope to reconstruct the return pointer from the stack pointer owned by the
+adversary. Instead, the old stack pointer should be stored as part of the
+activation record.
 
@@
 
-It's important that neither of these capabilities are accessible to the
-adversary. In other words, the return pointer provided to the adversary must be
-a capability that they can jump to but not read from, i.e.\ an enter
-capability. To make this work, we construct the activation record as depicted in
-Figure~\ref{fig:activ-rec-struct}. The return pointer is an $\entry$ capability
-with authority over the activation record containing the previous return and
-stack pointer as well as a number of restore instructions. The return pointer's
-current address points to those instructions, so that, upon invocation, they are
-executed and can load the old stack pointer and program counter back into the
-register file. Because the return pointer is given to the adversary as an enter
-pointer, they cannot get hold of the contents of the data structure, but after
-invocation, its permission is updated to $\exec$, so the activation record's
-contents become available to the restore instructions.
+Clearly, neither of these capabilities should be accessible by the adversary. In
+other words, the return pointer provided to the adversary must be a capability
+that they can jump to but not read from, i.e.\ an enter capability. To make this
+work, we construct the activation record as depicted in
+Figure~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
+over the entire activation record (containing the previous return and stack
+pointer), and its current address points to a number of restore instructions in
+the record, so that upon invocation, these instructions are executed and can
+load the old stack pointer and program counter back into the register file. As
+the return pointer is an enter pointer, the adversary cannot get hold of the
+activation record's contents, but after invocation, its permission is updated to
+$\exec$, so the contents become available to the restore instructions.
 
@@
 record? The attentive reader may already see that there is only one possibility:
-since the activation record contains the old stack pointer, which is necessarily
-local, the activation record can only be constructed in a part of memory where
-we have write-local access, i.e.\ on the stack. Note that this means we will be
-placing and executing instructions on the stack, i.e.\ it will not just contain
-code pointers and data. This means that our calling convention should be
-combined with protection against stack smashing attacks (i.e.\ buffer overflows
-on the stack overwriting activation records' contents), but conveniently, the
-fine-grained memory protection on capability machines is well-suited to easily
-provide such protection.
+since the activation record contains the old stack pointer, which is local, the
+activation record can only be constructed in a part of memory where we have
+write-local access, i.e.\ on the stack. Note that this means we will be placing
+and executing instructions on the stack, i.e.\ it will not just contain code
+pointers and data. This means that our calling convention should be combined
+with protection against stack smashing attacks (i.e.\ buffer overflows on the
+stack overwriting activation records' contents). Luckily, such protection is
+easy to provide thanks to the capability machine's fine-grained memory
+protection.
 
 \textbf{Invoking an adversary callback} If we have a higher-order interface to
-the adversary, we might need to invoke an adversary callback. In this case, not
-so much changes with respect to the situation where we invoke regular adversary
+the adversary, we may need to invoke an adversary callback. In this case, not so
+much changes with respect to the situation where we invoke static adversary
 code. The adversary can provide a callback as a capability for us to jump to,
-either an $\entry$-capability if they want to protect themselves from us or just an
-$\exec$ capability if they are not worried about that. However, there is one
-scenario that we need to prevent, namely that they construct the callback
-capability to point into the stack, where they may have stored local
-capabilities that they should not have access to upon invocation of the callback.
-Specifically, they might store return pointers and stack pointers from previous
-stack frames there, and illegally use them inside the callback.
+either an $\entry$-capability if they want to protect themselves from us or just
+an $\exec$ capability if they are not worried about that. However, there is one
+scenario that we need to prevent: if they construct the callback capability to
+point into the stack, it may contain local capabilities that they should not
+have access to upon invocation of the callback. As before, this includes return
+pointers and stack pointers from previous stack frames there, that they may be
+trying to illegally use inside the callback.
 
 To prevent this, we only accept callbacks from the adversary in the form of
-global capabilities and we check this prior to invoking them (and fail
-otherwise). This should not be an overly strict requirement, since our own
+global capabilities, which we dynamically check before invoking them (and we
+fail otherwise). This should not be an overly strict requirement: our own
 callbacks do not contain local data themselves, so there should be no need for
@@
 
-\textbf{Having a callback invoked by the adversary}
-The above leaves us with probably the hardest scenario: how to provide a
-callback to the adversary. The basic idea is that we allocate a block of memory
-using malloc that we fill with the capabilities and data that the callback needs
-access to, as well as some prelude instructions that load the data into
-registers and jumps to the right code. Note that this implies that no local
-capabilities can be stored as part of a closure. We can then provide the adversary with an
-enter-capability covering the allocated block and pointing to the contained
-prelude instructions. However, the question that remains in this setup is: where
-do we get the stack pointer?
+\textbf{Having a callback invoked by the adversary} The above leaves us with
+perhaps the hardest scenario: how to provide a callback to the adversary. The
+basic idea is that we allocate a block of memory using malloc that we fill with
+the capabilities and data that the callback needs access to, as well as some
+prelude instructions that load the data into registers and jumps to the right
+code. Note that this implies that no local capabilities can be stored as part of
+a closure. We can then provide the adversary with an enter-capability covering
+the allocated block and pointing to the contained prelude instructions. However,
+the question that remains in this setup is: where do we get a stack pointer when
+the callback is invoked?
 
 Our answer is that the adversary should provide it to us, much like we provide
-them with a current stack pointer when we invoke his code.  However, it is
-important that we do not just accept any capability as a stack pointer but check
-that it is safe to use.  Specifically, we check that it is indeed a $\rwlx$
-capability and make sure its current address points just below its start address.
-Without this check, an adversary could potentially get control over
-our local stack frame during a callback by passing us a local $\rwx$
-capability to a global data structure (i.e.\ there exists a global
-capability for it) as the stack pointer and a global
-return pointer. If our local state contains no local capabilities,
-then following our calling convention, the callback would not fail and
-the adversary could use a stored capability for the global data
-structure to access our local state. 
-To prevent this from happening, we need to make sure the stack
-capability is local and there exists no global capability for that
-memory segment. By making sure the capability has permission $\rwlx$,
-we rely on the system wide assumption that there are only local
-$\rwlx$ capabilities.
+them with a stack pointer when we invoke their code. However, it is important
+that we do not just accept any capability as a stack pointer but check that it
+is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability
+and make sure its current address points just below the start address. Without
+this check, an adversary could potentially get control over our local stack
+frame during a callback by passing us a local $\rwx$ capability to a global data
+structure (i.e.\ there exists a global capability for it) as the stack pointer
+and a global callback for our callback to invoke. If our local state contains no
+local capabilities, then following our calling convention, our callback would
+not fail and the adversary could use a stored capability for the global data
+structure to access our local state. To prevent this from happening, we need to
+make sure the stack capability is $\rwlx$, since the system wide assumption then
+tells us that the adversary cannot have a global capabilities to it.
 
@@
 \end{description}
-\textbf{Reusable macro instructions} 
-We define a number of reusable macros capturing the calling convention and other conveniences. The macros are:
+\textbf{Reusable macro instructions} We define a number of reusable macros
+capturing the calling convention and other conveniences. All macros that uses
+the stack assumes a stack pointer in register $r_\stk$.
 \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
-$\var{name}$ from the linking table and stores it in register
-$r$. \texttt{\footnotesize{push $r$}}/\texttt{\footnotesize{pop $r$}} works as one would expect. All
-macros that uses the stack assumes a stack pointer in register
-$r_\stk$.  \texttt{\footnotesize{prepstk $r$}} is used when a callback is invoked by
-the adversary and prepares the received stack pointer by checking that
-it has permission $\rwlx$ and making it conform to our calling convention. \texttt{\footnotesize{scall $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} Jumps
-to the capability in register $r$ in the manner described above. That
-is, it pushes local state (the contents of registers
-$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
-stack, sets up the necessary on stack return code and creates a return
-pointer, restricts the stack pointer, clears the unused part of the
-stack, clears the necessary registers and jumps to $r$. Upon return,
-the private state is restored. \texttt{\footnotesize{mclear $r$}} clears all the
-memory the capability in register $r$ has authority
-over. \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
+$\var{name}$ from the linking table and stores it in register $r$.
+\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
+would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
+invoked by the adversary and prepares the received stack pointer by checking
+that it has permission $\rwlx$ and making it conform to our calling convention.
+\texttt{\footnotesize{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
+capability in register $r$ in the manner described above. That is, it pushes
+local state (the contents of registers $\overline{r_{\var{priv}}}$), stack
+pointer and return pointer to the stack, sets up the necessary on stack return
+code and creates an $\entry$ return pointer, restricts the stack pointer, clears
+the unused part of the stack, clears the necessary registers and jumps to $r$.
+Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
+clears all the memory the capability in register $r$ has authority over.
+\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
 $\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
 register $r$ is a global capability. \texttt{\footnotesize{crtcls
-  $\overline{(x_i,r_i)}$ $r$}} allocates a closure pair where the
-capability in register $r$ is used to point to the program and a new
-environment is allocated (using malloc) where the contents of
-$\overline{r_i}$ is stored. In the program referred to by $r$, an
+    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
+closure's code and a new environment is allocated (using malloc) where the
+contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
 implicit fetch happens when an instruction refers to $x_i$.
@@
 
-%\lars{maybe we can present the examples informally before the logical
-%relation and then return to them after the logical relation (to
-%present the formal properties) ?}
-% \lau{not enough space to do this.}
-
 \section{Logical relation}
 \label{sec:logical-relation}
-In order to be able to reason about programs running on the capability
-machine and prove properties such as well-bracketedness and
-encapsulation, we define a unary logical relation.
+In order to reason about programs on the capability machine and prove properties
+such as well-bracketedness and encapsulation, we define a unary logical
+relation.
 % Intuition
-It is a semantic model of all the words on the capability machine that are
-definitely safe to give to an adversary. Whether a capability is safe to give to
-an adversary depends heavily on the contents of the memory since instructions as
-well as capabilities may reside in memory. Hence we index the logical relation
-over worlds, which are used to model the memory (making our logical relation a
-Kripke logical relation). A world consists of a finite set of regions, and each
-region can be understood as a protocol describing the possible evolution of part
-of the memory. Future evolutions of worlds (e.g. new protocols established over
-memory received from malloc, or steps in a protocol already established for a
-part of memory) are modeled using a future world relation. In fact, as we will
-explain below, there are two such relations to accomodate reasoning about local
-capabilities.
+It is a semantic model of all the words that are definitely safe to give to an
+adversary. Whether a capability is safe depends heavily on the contents of the
+memory since instructions as well as capabilities may reside there. Hence we
+index the logical relation over worlds, modelling the memory (this makes our
+logical relation into a Kripke logical relation). A world consists of a finite
+set of regions, and each region defines a protocol that governs the possible
+evolution of a part of memory. A future world relation defines how worlds may
+evolve during execution (e.g. new protocols established over memory received
+from malloc, or steps in a protocol already established for a part of memory).
+In fact, as we will explain below, there are two future world relations to
+accomodate reasoning about local capabilities.
 %
@@
 Each permanent and temporary region contains a state transition system, with
-public and private transitions, which describes the protocol governing
-the memory described by the region, similarly to what
-has been used in logical relations for high-level
-languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
-The protocols imposed on memory by permanent regions stay in place
-indefinitely. This means that any capability, local or global, can
-depend on these protocols. The protocols imposed on memory by temporary
-regions can be revoked at certain points during execution,
-corresponding to the points, on the condition that we clear local capabilities can be
-revoked. This means that local capabilities safely can depend on the
-protocols imposed on memory by temporary regions, whereas global
-capabilities cannot, since a global capability may stay around
-throughout the execution of a program, but a temporary region may be
-revoked. This is illustrated in \figurename~\ref{fig:cap-world}.
-
-For technical reasons, a temporary region is not removed in the model when it is
-revoked. Instead it is turned into a new kind of region, namely a revoked
-region. A revoked region contains no state transition system and puts no
-requirements on the memory it models. It simply serves as a mask for a
-temporary region that has been revoked. This way of masking a region goes back
-to earlier work of \citet{Ahmed2004semantics} and was also used
-by \citet{Thamsborg:2011:KLR:2034773.2034831}.
+public and private transitions, to describe a protocol for the region's memory,
+similarly to what has been used in logical relations for high-level
+languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
+imposed by permanent regions stay in place indefinitely. Any capability, local
+or global, can depend on these protocols. Protocols imposed by temporary regions
+can be revoked at certain points during execution, on the condition that we
+clear local capabilities when this happens. This means that local capabilities
+can safely depend on the protocols imposed by temporary regions, but global
+capabilities cannot, since a global capability may outlive a temporary region
+that is revoked. This is illustrated in \figurename~\ref{fig:cap-world}.
+
+For technical reasons, we do not actually remove a revoked temporary region from
+the world, but we turn it a special revoked region that exists for this purpose.
+Such a revoked region contains no state transition system and puts no
+requirements on the memory. It simply serves as a mask for a revoked temporary
+region. Masking a region like this goes back to earlier work of
+\citet{Ahmed2004semantics} and was also used by
+\citet{Thamsborg:2011:KLR:2034773.2034831}.
 
@@
 \emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
-correctness, we do not prove a general theorem that shows this, because it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
+correctness, we do not prove a general theorem that shows this, it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
@@
 as the code returns to a stable state (i.e. transitions to a public future world
-of the original) before returning (i.e.
-invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation).
-Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
-We
+of the original) before returning (i.e. invoking the continuation). We
 generalise this idea to reason about local capabilities: validity of local

### 2017-02-15 csf-2017/paper.tex
commands to disable comments
Add some more explanation about the world existence theorem.

@@
 %\dominique{``end up being defined recursively'' does not sound very insightful}
-The worlds are defined recursively: Worlds describe memory, which can
-contain capabilities, whose safety depends on worlds. We use the method of
-\citet{Birkedal:2011:SKM:1926385.1926401} for constructing the worlds.
-
+A region defines what its part of memory can legally look like. However,
+validity of the memory's contents may itself depend on the state of other parts
+of memory. To accomodate this, worlds need to be defined recursively: Worlds
+describe memory, which can contain capabilities, whose safety is
+world-dependent. Recursive worlds are a common requirement in Kripke models and
+the following lemma uses the method of \citet{Birkedal:2011:SKM:1926385.1926401}
+for constructing them. The formulation of the lemma is technical, so we
+recommend that non-expert readers ignore the technicalities and accept that
+there exists a set of worlds $\Wor$ and two relations $\futurestr$ and
+$\futurewk$ satisfying the (recursive) equations in the theorem (where the
+$\blater$ operator can be safely ignored).
 % Short technical explanation
@@
   \includegraphics{w11}
-  \caption{Illustration of the relation between local and global capabilities and temporary and permanent regions.
-The table represents a memory and the colored fields are regions governing parts of the memory.
-Roughly speaking, global capabilities should never depend on memory governed by temporary regions.}
+  \caption{An illustration of local/global capabilities and temporary/permanent
+    regions. The table represents a memory and the colored fields are regions
+    governing a memory range. Global capabilities cannot depend on temporary
+    regions.}
   \label{fig:cap-world}

### 2017-02-15 csf-2017/paper.tex
More shortening.

@@
 \subsubsection{Future worlds}
-The future world relations model how memory may evolve over time. 
-The \emph{public future world} 
-$W' \futurewk W$ is defined to mean that $\dom(W') \supseteq \dom(W)$
-and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$.  That is,
-in a public future world, more regions may have been allocated, and
-each of the regions may also have evolved according to the public future
-region relation (defined below). The \emph{private future world} relation
-$W' \futurestr W$ is defined similarly, using a private future region
-relation.
-
-The properties of the \emph{private future region} relation are:
-\begin{mathpar}
-  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
-              (s,s') \in \phi} 
-  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ r \in \Regions }
-  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ r \in \Regions }
-  { r \futurestr \revoked }
-\end{mathpar}
-These rules express that, in the private future region relation, we allow
-revocation of temporary regions. As explained above, this can be done using a
-special revoked region, which does not describe any memory, but we also allow
-any other region to serve as a mask, and thus any region $r$ can be a future
-region of a temporary region and of a revoked region. On the other hand,
-permanent regions cannot be masked away. They are only allowed to take a
-transition in the private part of the transition system.
-
-The \emph{public future} region relation satisfies
+The future world relations model how memory may evolve over time. The
+\emph{public future world} $W' \futurewk W$ requires that $\dom(W') \supseteq
+\dom(W)$ and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$. That is, in a
+public future world, new regions may have been allocated, and existing regions
+may have evolved according to the public future region relation (defined below).
+The \emph{private future world} relation $W' \futurestr W$ is defined similarly,
+using a private future region relation.
+
+The \emph{public future} region relation is the simplest. It satisfies
 % \dominique{is defined by?} 
@@
 \end{mathpar}
-In the public future regions we do not permit revocation of temporary regions.
-Instead, temporary and permanent regions are only allowed to take a transition
-in the public part of the transition system. Revoked regions can either remain
-revoked or be masked away by a temporary region. This means that the public
-future world relations allows us to reuse a region that has been revoked
+Both temporary and permanent regions are only allowed to transition according to
+the public part of their transition system. Additionally, revoked regions must
+either remain revoked or be replaced by a temporary region. This means that the
+public future world relations allows us to reuse a region that has been revoked
 earlier.
+
+The \emph{private future region} relation satisfies:
+\begin{mathpar}
+  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
+              (s,s') \in \phi} 
+  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ r \in \Regions }
+  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
+  \and
+  \inferrule{ r \in \Regions }
+  { r \futurestr \revoked }
+\end{mathpar}
+Here, revocation of temporary regions is allowed. As explained above, this can
+be done by replacing them with a special revoked region, which does not describe
+any memory, but in fact, any other region can also be used to replace both
+revoked and temporary regions. On the other hand, permanent regions cannot be
+masked away. They are only allowed to transition according to the private part
+of the transition system.
 

### 2017-02-15 csf-2017/paper.tex
Small modification before I head home.

@@
 \subsubsection{World satisfaction}
-A memory satisfies a world, written $\memSat{\ms}{W}$, 
-if it can be partitioned into disjoint parts such that each part is
-in the interpretation of an active (permanent or temporary) region. 
+A memory satisfies a world, written $\memSat{\ms}{W}$, if it can be partitioned
+into disjoint parts such that each part is accepted by an active (permanent or
+temporary) region.
 \begin{align*}
@@
 %   respect them itself too.} 
-The logical relation defines semantically what it
-means to be capability safe.
-% \dominique{We have not introduced capability
-%   safety..} 
-In large parts, capability safety is defined by when a capability
-can be safely given to an adveserial piece of code. In the following paragraphs,
-we will go through the definitions of the logical relation and provide an
-intuition why they define capability safety. The definition of the logical
-relation is found in \figurename~\ref{fig:logrel}, and \figurename~\ref{fig:perm-conds}.
-Some definitions are omitted and explained when needed. The precise definitions
-can be found in the technical appendix~\cite{technical_appendix}.
+The logical relation now defines semantically what it means to be capability
+safe. In large parts, capability safety is defined by when a capability can be
+safely given to adversarial code. In the following paragraphs, we will go
+through the definitions of the logical relation and provide an intuition why
+they define capability safety. The definition of the logical relation is found
+in \figurename~\ref{fig:logrel}, and \figurename~\ref{fig:perm-conds}. Some
+definitions are omitted and explained when needed. The precise definitions can
+be found in the technical appendix~\cite{technical_appendix}.
 

### 2017-02-15 csf-2017/paper.tex
Spell checking

@@
 \newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+
+% Uncomment me to disable comments
+\renewcommand\lau[1]{}
+\renewcommand\dominique[1]{}
+\renewcommand\lars[1]{}
 
@@
 stored to memory for which we have a capability with write-local authority
-(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accomodate a stack,
+(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
 where register contents can be stored, including local capabilities. As long as all
@@
 default implementation of the CCall exception handler forbids local capabilities
-from being passed accross module boundaries. Such a restriction fundamentally
+from being passed across module boundaries. Such a restriction fundamentally
 breaks our calling convention, since we pass around local return pointers and
@@
 
-\textbf{Program startup} We assume that the language runtime initializes the
+\textbf{Program start-up} We assume that the language runtime initializes the
 memory as follows: a contiguous array of memory is reserved for the stack, for
@@
 instructions or static data. Another part of memory (called the heap) is
-initially governed by malloc and at program startup, no other code has
+initially governed by malloc and at program start-up, no other code has
 capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
@@
 \begin{description}[font=\normalfont\itshape]
-\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
+\item[At program start-up] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
 \item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
@@
 In fact, as we will explain below, there are two future world relations to
-accomodate reasoning about local capabilities.
+accommodate reasoning about local capabilities.
 %
@@
 validity of the memory's contents may itself depend on the state of other parts
-of memory. To accomodate this, worlds need to be defined recursively: Worlds
+of memory. To accommodate this, worlds need to be defined recursively: Worlds
 describe memory, which can contain capabilities, whose safety is
@@
 The Fundamental Theorem allows us to prove capability safety for
-adveserial programs: if the authority granted by
-the capabilty for the untrustred program is reasonable, then it is safe to
+adversarial programs: if the authority granted by
+the capability for the adversarial program is reasonable, then it is safe to
 execute. Intuitively, this is because our logical relation captures the
@@
 that the adversary capability is an enter capability. Luckily the enter
-capability becomes $\exec$ after the jump and then, the Fundamental Theorem
+capability becomes $\exec$ after the jump and then the Fundamental Theorem
 applies.
@@
 
-In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accomodate this in the
+In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in the
 proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
@@
 capabilities can be seen as a generalisation of the syntactically-enforced
-unstoreable status of evaluation contexts in lambda calculi without control
+unstorable status of evaluation contexts in lambda calculi without control
 effects (of which well-bracketed control flow is a consequence). To reason about
@@
 formalising the idea that it is safe to make temporary internal state
-modifications (private world transititions, that invalidate the continuation,
+modifications (private world transitions, that invalidate the continuation,
 but not other values) while an expression is performing internal steps, as long

### 2017-02-15 csf-2017/paper.tex
Style corrections

@@
   enforce properties like control-flow correctness, encapsulation of local
-  state. We provide a formalisation of a representative capability machine and
+  state. We provide a formalization of a representative capability machine and
   propose a novel calling convention for enforcing control-flow correctness and
@@
 encapsulation of private state in ML-like languages). And even if such
-a scheme were defined, a formal proof depends on a formalisation of
+a scheme were defined, a formal proof depends on a formalization of
 the encapsulation provided by the capability machine at hand.
@@
 execute. Intuitively, this is because our logical relation captures the
-safe behaviours of the capability machine that do not allow you to
+safe behaviors of the capability machine that do not allow you to
 produce something \emph{unsafe} from something safe. 
@@
 future worlds. The main insight is that the temporary, revokable nature of these
-capabilities can be seen as a generalisation of the syntactically-enforced
+capabilities can be seen as a generalization of the syntactically-enforced
 unstorable status of evaluation contexts in lambda calculi without control
@@
 \citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
-of previous work. The capability machine formalised in
+of previous work. The capability machine formalized in
 Section~\ref{sec:capab-mach-with} is a simple but representative
@@
 every indirect jump and using memory access control to prevent overwriting code,
-this work enforces what they call control-flow integrity, formalised as the
+this work enforces what they call control-flow integrity, formalized as the
 property that every jump will follow a legal path in the control-flow graph.
@@
 with a primitive notion of compartments) and we do not use a trusted stack
-manager component, but a decentralised calling convention based on local
+manager component, but a decentralized calling convention based on local
 capabilities. Also, both prove a secure compilation result from a high-level

### 2017-02-15 csf-2017/paper.tex
more style changes

@@
 language properties relied upon by software authors, such as control-flow
-correctness, local state encapsulation, etc. Commodity processors offer little
+correctness, local-state encapsulation, etc. Commodity processors offer little
 support for defending against such attacks: they offer security primitives with
@@
 scalability. As a result, defenses against attacks on control-flow correctness
-and local state encapsulation are either limited to only certain common forms of
+and local-state encapsulation are either limited to only certain common forms of
 attacks (leading to an attack-defense arms race) or rely on techniques like
@@
 primitives on capability machines lend themselves well to the
-enforcement of properties like local state encapsulation and
+enforcement of properties like local-state encapsulation and
 control-flow correctness. Potentially, they will enable new compilation
@@
 \item Demonstrate our results by applying them to challenging examples,
-  specifically constructed to demonstrate local state encapsulation and
+  specifically constructed to demonstrate local-state encapsulation and
   control-flow correctness guarantees in the presence of cross-component
@@
 pointer, it could access the private data stored there by the trusted code, and
-break local state encapsulation.
+break local-state encapsulation.
 
@@
 In this section, we demonstrate how our formal model of capability
-safety allows us to prove local state encapsulation and control-flow
+safety allows us to prove local-state encapsulation and control-flow
 correctness properties for challenging program examples. The security

### 2017-02-15 csf-2017/paper.tex
a few changes in introduction and capability machine section (taking a break until later/tomorrow)

@@
 enforce a strict separation between numbers and capabilities, a kind
-of pointers carrying some kind of authority. Memory capabilities carry
+of pointers carrying authority. Memory capabilities carry
 the authority to read and/or write to a range of memory
@@
 both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
-CHERI~\cite{Watson2015Cheri}. To avoid uninteresting detail, we assume an
+CHERI~\cite{Watson2015Cheri}. To avoid uninteresting details, we assume an
 infinite address space and unbounded integers.
@@
 authority to execute permissions $\perm$ on the memory range
-$[\var{base},\var{end}]$, together with a current address $\addr$ and a tag
+$[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag
 $\gl$ indicating whether the capability is global or local. There is no notion
@@
 modifying capabilities: $\lea{}{}$ (modifies the current address),
-$\restricttwo{}{}$ (modifies the permission and local/global flag) and
+$\restricttwo{}{}$ (modifies the permission and local/global tag) and
 $\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
@@
 number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and
-$\geta{}{}$ provide access to a capability's permissions, local/global flag, base,
+$\geta{}{}$ provide access to a capability's permissions, local/global tag, base,
 end and current address, respectively.
@@
 number of representative instructions. Essentially, a configuration $\Phi$ will
-either decode and execute the instruction at $\memreg{\pcreg}$ if it is
+either decode and execute the instruction at $\memreg(\pcreg)$ if it is
 executable and the current address is in the valid range or fail otherwise. The
@@
 authority (i.e.\ read permission and an appropriate range). The $\restricttwo{}{}$
-instruction will update a capability's permissions and global/local status in
+instruction will update a capability's permissions and global/local tag in
 the register file, but requires that the new permissions are weaker than the
@@
 $\pcreg$\footnote{Note that the capability in the $\pcreg$-register can be used
-  to load from.}.
+  to load from.}.% Lau, du er nået hertil. Husk stak vs code pointe.
 % \lau{As we

### 2017-02-15 csf-2017/paper.tex
Slight tuning of the abstract.

@@
   Capability machines provide security guarantees at machine level which makes
-  them an interesting target for compilation schemes that aim to provably
-  enforce properties like control-flow correctness, encapsulation of local
-  state. We provide a formalization of a representative capability machine and
-  propose a novel calling convention for enforcing control-flow correctness and
-  encapsulation of local state on it. To prove these properties, we provide a
-  logical relation that semantically captures the guarantees provided by the
-  hardware (a form of capability safety). These results are not tied to our
-  calling convention and can be used to reason about arbitrary programs.
+  them an interesting target for compilation schemes that provably enforce
+  properties like control-flow correctness, encapsulation of local state. We
+  provide a formalization of a representative capability machine with local
+  capabilities and propose a novel calling convention for enforcing control-flow
+  correctness and encapsulation of local state on it. To prove these properties,
+  we provide a logical relation that semantically captures the guarantees
+  provided by the hardware (a form of capability safety). These results are not
+  tied to our calling convention and can be used to reason about arbitrary
+  programs.
 \end{abstract}

### 2017-02-15 csf-2017/paper.tex
updates up to LR section

@@
 $\pcreg$\footnote{Note that the capability in the $\pcreg$-register can be used
-  to load from.}.% Lau, du er nået hertil. Husk stak vs code pointe.
+  to load from.}.
 % \lau{As we
@@
 \emph{temporarily} given to an adversary, for the duration of an invocation: if
-we take care to erase the capability from the register file after control is
+we take care to clear the capability from the register file after control is
 passed back to us, they will not have been able to store the capability.
@@
 for various purposes (register spilling, storing local state when invoking other
-functions etc.), so we also need to erase that data before returning to the
+functions etc.), so we also need to clear that data before returning to the
 adversary.
@@
 capability. If it were global, the adversary would be able to store away the
-return pointer in a global data structure, and jump to it later, in
+return pointer in a global data structure (i.e.\ there exists a global capability for it), and jump to it later, in
 circumstances where this should not be possible. For example, they could store
@@
 have access to upon invocation of the callback. As before, this includes return
-pointers and stack pointers from previous stack frames there, that they may be
+pointers and stack pointers from previous stack frames that they may be
 trying to illegally use inside the callback.
@@
 perhaps the hardest scenario: how to provide a callback to the adversary. The
-basic idea is that we allocate a block of memory using malloc that we fill with
+basic idea is that we allocate blocks of memory using malloc that we fill with
 the capabilities and data that the callback needs access to, as well as some
@@
 that we do not just accept any capability as a stack pointer but check that it
-is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability
-and make sure its current address points just below the start address. Without
+is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability. Without
 this check, an adversary could potentially get control over our local stack
 frame during a callback by passing us a local $\rwx$ capability to a global data
-structure (i.e.\ there exists a global capability for it) as the stack pointer
+structure instead of a proper stack pointer
 and a global callback for our callback to invoke. If our local state contains no
-local capabilities, then following our calling convention, our callback would
+local capabilities, then, following our calling convention, the callback would
 not fail and the adversary could use a stored capability for the global data
 structure to access our local state. To prevent this from happening, we need to
-make sure the stack capability is $\rwlx$, since the system wide assumption then
-tells us that the adversary cannot have a global capabilities to it.
+make sure the stack capability gives $\rwlx$ authority, since the system wide assumption then
+tells us that the adversary cannot have a global capabilities to our local stack.
 
@@
 invoked by the adversary and prepares the received stack pointer by checking
-that it has permission $\rwlx$ and making it conform to our calling convention.
+that it has permission $\rwlx$.
 \texttt{\footnotesize{scall
@@
 the stack that we have used ourselves, but we need to clear all of the stack that we are not using ourselves. 
-In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be erased.
+In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be cleared.
 We believe this is a common requirement for typical usage scenarios of local capabilities and

### 2017-02-16 csf-2017/paper.tex
g1 comment about optimization to g1 calling convention
rewrote a sentence in the introduction

@@
 relies on \emph{local capabilities}, a type of capabilities offered by CHERI
-that can be passed over a function invocation \emph{temporarily}, for the
-duration of the invocation, and revoked afterwards. However, a lot of details
+to \emph{temporarily} relinquish authority, namely for the duration of a function invocation whereafter the capability can be revoked.
+However, a lot of details
 remain to be investigated (e.g., how to combine a trusted shadow stack with
@@
 generates closures with one variable $x$ set to $0$ in its environment and
-\texttt{\footnotesize{f4}} as the program. \texttt{\footnotesize{f4}} expects one argument, a callback. The
+\texttt{\footnotesize{f4}} as the program (note that we here use knowledge about the stack usage to make optimizations to the calling convention).
+\texttt{\footnotesize{f4}} expects one argument, a callback.
+The
 program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}

### 2017-02-16 csf-2017/paper.tex
moved figure 5 to help latex not waste space
A bit of rewording.  Don't use capability safety as a property of values or terms, as it is generally used as a property of the language rather than values or terms.

--- 2017-02-15 csf-2017/paper.tex
@@
 %   respect them itself too.} 
-The logical relation now defines semantically what it means to be capability
-safe. In large parts, capability safety is defined by when a capability can be
-safely given to adversarial code. In the following paragraphs, we will go
-through the definitions of the logical relation and provide an intuition why
-they define capability safety. The definition of the logical relation is found
-in \figurename~\ref{fig:logrel}, and \figurename~\ref{fig:perm-conds}. Some
-definitions are omitted and explained when needed. The precise definitions can
-be found in the technical appendix~\cite{technical_appendix}.
-
-The \emph{observation relation} $\observations$ defines which configurations
-are capability safe. Roughly speaking, a configuration is capability safe if,
-whenever it halts, then the resulting memory satisfies a private future world.
-Notice that failing is considered safe behavior and in fact, the machine will
-often resort to failing when, for example, an unauthorized
-access is attempted, like a load of a memory location through a
-capability without read permission. This is similar to the logical relation
-given for an untyped language in~\cite{Devriese:2016ObjCap}, but unlike logical
-relations for typed languages, which typically require that computations do not
-fail.
+The logical relation now defines semantically when values, terms, and
+configurations are for execution by the adversary, i.e. without the risk that
+they will be able to use it to break protocols in the world. The definition of
+the logical relation is found in \figurename{}s~\ref{fig:logrel}
+and~\ref{fig:perm-conds} and we provide some explanations in the following
+paragraphs. We omit some definitions and explain them only verbally, but precise
+definitions can be found in the technical appendix~\cite{technical_appendix}.
+
+The \emph{observation relation} $\observations$ defines configurations that we
+consider safe. Roughly speaking, a configuration is safe if, whenever it halts,
+there is a private future world that the resulting memory satisfies. Notice that
+failing is considered safe behavior. In fact, the machine often resorts to
+failing when an unauthorized access is attempted, like loading from a capability
+without read permission. This is similar to \cite{Devriese:2016ObjCap}'s logical
+relation for an untyped language, but unlike typical logical relations
+for typed languages, which require that programs do not fail.
 
@@
 safe when they contain safe words (i.e.\ words in $\stdvr$) in all registers but
-$\pcreg$. The \emph{expression relation} $\stder$ defines when a word is safe
-to use as a program counter. That is the case when the word can be plugged into
-a capability safe register file (i.e.\ a register file in $\stdrr$) and paired
-with a memory satisfying the world to become a capability safe configuration.
+$\pcreg$. The \emph{expression relation} $\stder$ defines that a word is safe to
+use as a program counter if it can be plugged into a safe register file (i.e.\ a
+register file in $\stdrr$) and paired with a memory satisfying the world to
+produce a safe configuration.
 
@@
 to manipulate memory and potentially break memory protocols, so they need to
-satisfy certain conditions to be capability safe. In
-\figurename~\ref{fig:perm-conds}, we define such a condition for each kind of
-permission a capability can have (we will call these permission-based
-conditions).
+satisfy certain conditions to be safe. In \figurename~\ref{fig:perm-conds}, we
+define such a condition for each kind of permission a capability can have (we
+will call these permission-based conditions).
 
 For capabilities with read permission, the $\readCond{}$ ensures that it can
-only be used to read capability safe words, by requiring that the region
-governing the relevant part of memory, guarantees that the memory contains only
-words in the value relation. Generally speaking, global capabilities should only
-depend on permanent regions. Therefore we use the function
-$\var{localityReg}(\gl,W)$, which projects out all active regions when
-$\gl$ (the locality) is local, but only the permanent regions when $\gl$ is global.
-The read condition is defined in terms of a standard region
-$\iota^\pwl_{\start,\addrend}$ which requires all the words in the range
-$[\start,\addrend]$ to be safe.
-
-For a capability with write permission, $\writeCond{}$ must be
-satisfied for the capability's range of authority. Such a capability
-may be used to write any word, so the region governing the relevant
-part of memory must (at least) allow any capability safe word to be
-stored. However, this is not entirely true because of the distinction
-between write and permit-write-local capabilities.  Only
-permit-write-local capabilities need to allow arbitrary safe values to
-be stored, while for write capabilities, it is sufficient to allow
-storing values that are both non-local and safe. The write condition
-uses $\var{localityReg}$ to pick an appropriate region for its
-locality in the same way the read condition does. The condition for
-the region to be address-stratified is there for technical reasons and for the full definition, we refer to the technical appendix. In
-the value relation, the write condition is used with $\iota^\pwl$ for
-write-local capabilities. For write capabilities, the write condition
-is used with another standard region $\iota^\nwl_{\start,\addrend}$
-which requires all the words in the range $[\start,\addrend]$ to be
-non-local and safe.
-
-The conditions $\entryCond{}$ and $\execCond{}$ are very similar and differ only
-slightly. In both cases, we require that the capability can be safely jumped to.
-However, executable capabilities can be updated to point anywhere in memory, so
-they must be capability safe as a program counter (in the $\stder$-relation) no
-matter where they point to. In contrast, enter capabilities are opaque and can
-only be used to jump to the address they point to. They also change permission
-when jumped to, so we require them to be capability safe as a program counter
-when the permission is changed to read/execute. Because the capabilities are not
-necessarily invoked immediately, we need to require that this is the case in a
-future world, but it depends on the locality of the capability which future
-worlds we consider. If the capability is global, then it should only depend on
-permanent regions, so we require capability safety as a program counter in any
-\emph{private} future world (recall that temporary regions may be revoked in a
-private future world). For local capabilities, it suffices to be capability safe
-in \emph{public} future worlds, where temporary regions are still there.
+only be used to read safe words, by requiring that the region governing the
+relevant part of memory, guarantees that it contains only words in the value
+relation. Generally speaking, global capabilities should only depend on
+permanent regions. Therefore we use the function $\var{localityReg}(\gl,W)$,
+which projects out all active regions when $\gl$ (the locality) is local, but
+only the permanent regions when $\gl$ is global. The read condition is defined
+in terms of a standard region $\iota^\pwl_{\start,\addrend}$ which requires all
+the words in the range $[\start,\addrend]$ to be safe.
+
+For a capability with write permission, $\writeCond{}$ must be satisfied for the
+capability's range of authority. Such a capability may be used to write any
+word, so the region governing the relevant memory must (at least) allow any safe
+word to be stored. More precisely, for permit-write-local capabilities it needs
+to allow arbitrary safe values to be stored, while for write capabilities,
+storing values that are both non-local and safe is sufficient. Like before, we
+use $\var{localityReg}$ to pick an appropriate region for the capability's
+locality. The condition for the region to be address-stratified is a technical
+requirement, defined in the technical appendix. In the value relation, the write
+condition is used with $\iota^\pwl$ for write-local capabilities. For write
+capabilities, it is used with another standard region
+$\iota^\nwl_{\start,\addrend}$ which requires that the words in
+$[\start,\addrend]$ are non-local and safe.
+
+The conditions $\entryCond{}$ and $\execCond{}$ are very similar. Both require
+that the capability can be safely jumped to. However, executable capabilities
+can be updated to point anywhere in their range, so they must be safe as a
+program counter (in the $\stder$-relation) no matter where they point to. In
+contrast, enter capabilities are opaque and can only be used to jump to the
+address they point to. They also change permission when jumped to, so we require
+them to be safe as a program counter when the permission is changed to $\exec$.
+Because the capabilities are not necessarily invoked immediately, this must be
+true in any future world, but it depends on the locality of the capability which
+future worlds we consider. If it is global, then it should only depend on
+permanent regions, so we require safety as a program counter in \emph{private}
+future worlds (recall that temporary regions may be revoked in a private future
+world). For local capabilities, it suffices to be safe in \emph{public} future
+worlds, where temporary regions are still present.
 

### 2017-02-16 csf-2017/paper.tex

@@
 \subsection{Worlds}
+\begin{figure}
+  \centering
+  \includegraphics{w11}
+  \caption{An illustration of local/global capabilities and temporary/permanent
+    regions. The table represents a memory and the colored fields are regions
+    governing a memory range. Global capabilities cannot depend on temporary
+    regions.}
+  \label{fig:cap-world}
+\end{figure}
 % \dominique{Perhaps we could first try to convey the intuition of a permanent
@@
   \begin{align*}
-      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions) \\
+      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions)\\
       \Regions & = \{\revoked\} \uplus \\
@@
 \]
-
-\begin{figure}
-  \centering
-  \includegraphics{w11}
-  \caption{An illustration of local/global capabilities and temporary/permanent
-    regions. The table represents a memory and the colored fields are regions
-    governing a memory range. Global capabilities cannot depend on temporary
-    regions.}
-  \label{fig:cap-world}
-\end{figure}
 

### 2017-02-16 csf-2017/paper.tex
Add some more explanation about the fundamental theorem.

@@
 \end{theorem}
-The Fundamental Theorem allows us to prove capability safety for
-adversarial programs: if the authority granted by
-the capability for the adversarial program is reasonable, then it is safe to
-execute. Intuitively, this is because our logical relation captures the
-safe behaviors of the capability machine that do not allow you to
-produce something \emph{unsafe} from something safe. 
-The Fundamental Theorem also serves as a sanity check of the
-logical relation as it guarantees that a large class of capabilities
-actually inhabits it.
+The Fundamental Theorem can be understood as a general expression of the
+guarantees offered by the capability machine, an instance of a general property
+called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}. To
+understand this, consider that the theorem says that the capability
+$((\perm,\gl),\start,\addrend,\addr)$ is safe as a program counter, without any
+assumption about what instructions it actually points to (the only assumptions
+we have are about the read or write authority that it carries). As such, the
+theorem expresses the capability safety of the machine, which guarantees that
+\emph{any} instruction is fine and will not be able to go beyond the authority
+of the values it has access to.
+
+From another point of view, the Fundamental Theorem also serves as a sanity
+check of the logical relation as it guarantees that a large class of
+capabilities actually inhabits it.
 

### 2017-02-16 csf-2017/paper.tex
Shorten.

@@
 \label{sec:examples}
-In this section, we demonstrate how our formal model of capability
-safety allows us to prove local-state encapsulation and control-flow
-correctness properties for challenging program examples. The security
-measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
-ensure these properties. Since we are dealing with assembly language,
-there are many details to the formal treatment, and therefore we
-necessarily have to omit some details in the lemmas we state here.
-The interested reader can find all the technical details in the
-technical appendix~\cite{technical_appendix}.
+In this section, we demonstrate how our formalisation of capability safety
+allows us to prove local-state encapsulation and control-flow correctness
+properties for challenging program examples. The security measures of
+Section~\ref{sec:stack-and-return-pointer} are deployed to ensure these
+properties. Since we are dealing with assembly language, there are many details
+to the formal treatment, and therefore we necessarily have to omit some details
+in the lemmas we state here. The interested reader can find all the technical
+details in the technical appendix~\cite{technical_appendix}.
 
 \subsection{Encapsulation of local state}
-The capability machine provides encapsulation of local state. The programs
 \texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
-\figurename~\ref{fig:prog-f1-and-f2} demonstrate this. They are very similar: both
-store some local state, call an untrusted piece of code ($\adv$), and then test
-whether the local state is unchanged. They differ in the way they do this.
-Program \texttt{\footnotesize{f1}} assumes it has access to a stack pointer and uses the stack
-to store its local state. It also uses the stack-based calling convention
-\texttt{\footnotesize{scall}} to call the adversary. On the other hand, \texttt{\footnotesize{f2}} uses
-malloc to allocate memory for its local state and uses an activation-record
-based calling convention (described in the technical appendix) to run the
-adversarial code.
+\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's
+encapsulation of local state. They are very similar: both store some local
+state, call an untrusted piece of code ($\adv$), and then test whether the local
+state is unchanged. They differ in the way they do this. Program
+\texttt{\footnotesize{f1}} assumes access to a stack pointer and uses the stack
+to store its local state and the stack-based calling convention
+\texttt{\footnotesize{scall}} to call the adversary. On the other hand,
+\texttt{\footnotesize{f2}} uses malloc to allocate memory for its local state
+and uses an activation-record based calling convention (described in the
+technical appendix) to run the adversarial code.
 
@@
   \end{minipage}
-  \caption{Two simple programs that rely on local-state encapsulation. \texttt{f1} uses our stack-based calling convention. \texttt{f2} uses a calling not reliant on the stack}
+  \caption{Two example programs that rely on local-state encapsulation. \texttt{f1} uses our stack-based calling convention. \texttt{f2} does not rely on a stack.}
   \label{fig:prog-f1-and-f2}
@@
 
-For both programs, we can prove that if they are linked with an
-adversary, $\adv$, that is allowed to allocate new memory but
-otherwise has no capabilities, then the assertion will never fail when
-executing the program (see Lemmas~\ref{lem:correctness-f1}
-and~\ref{lem:correctness-f2} below).
+For both programs, we can prove that if they are linked with an adversary,
+$\adv$, that is allowed to allocate memory but has no other capabilities, then
+the assertion will never fail during executing (see
+Lemmas~\ref{lem:correctness-f1} and~\ref{lem:correctness-f2} below).
 % \lars{shouldn't we say already
@@
 
-In order to formulate lemmas about \texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}}, we
-need a way to observe whether the assertion fails. To this end, we
-assume access to a flag (an address in memory). When the assertion
-fails, it sets the flag to $1$ and halts. A slightly simplified
-version of the correctness lemma for \texttt{\footnotesize{f1}} is:
+In order to formulate results about \texttt{\footnotesize{f1}} and
+\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
+fails. To this end, we assume they have access to a flag (an address in memory)
+that they set to $1$ if the assertion fails and then halt. The correctness lemma
+for \texttt{\footnotesize{f1}} then states:
 \begin{lemma}
@@
 \]
-  where each of the capabilities have an appropriate range of authority and pointer. Furthermore
+where each of the capabilities have an appropriate range of authority and
+pointer\footnote{These assumptions are kept intentionally vague for brevity.
+  Full statements are in the technical appendix~\citep{technical_appendix}.}.
+Furthermore
   \begin{itemize}
@@
 To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
-configuration is capability safe (in the $\observations$ relation) for a world
-with a permanent region that requires the assertion flag to be 0. We prove this
-using an anti-reduction lemma by which it suffices to show that the
-configuration is capability safe after some reduction steps. We then use a general lemma
-for reasoning about \texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
-configuration that \texttt{\footnotesize{scall}} will jump to is capability safe and (2) that
-the configuration just after \texttt{\footnotesize{scall}} is done cleaning up is capability
-safe. We use the Fundamental Theorem to reason about the unknown code, but
-notice that the Fundamental Theorem says nothing about enter capabilities and
-that the adversary capability is an enter capability. Luckily the enter
-capability becomes $\exec$ after the jump and then the Fundamental Theorem
-applies.
+configuration is safe (in the $\observations$ relation) for a world with a
+permanent region that requires the assertion flag to be 0. By an anti-reduction
+lemma, it suffices to show that the configuration is safe after some reduction
+steps. We then use a general lemma for reasoning about
+\texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
+configuration that \texttt{\footnotesize{scall}} will jump to is safe and (2)
+that the configuration just after \texttt{\footnotesize{scall}} is done cleaning
+up is safe. We use the Fundamental Theorem to reason about the unknown code, but
+notice that the adversary capability is an enter capability, which the
+Fundamental Theorem says nothing about. Luckily the enter capability becomes
+$\exec$ after the jump and then the Fundamental Theorem applies.
 
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer
+  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer,
   if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
@@
 
-In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order for the
-assertion in the middle to succeed, they need to be well-bracketed. If the
-adversary was able to store the return pointer from the first call and invoke
-it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on top of its stack and
-the assertion will fail. However, the security measures of our calling
-convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
-sort of attack: specifically, the return pointer is local, so it can only be
-stored on the stack, but the part of the stack that is accessible to the
-adversary is cleared before the second invocation. The following lemma shows
-that there are also no other attacks that will break well-bracketedness of this
-example, i.e.\ the assertion never fails. It is similar to the two previous
-lemmas:
+In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order
+for the assertion in the middle to succeed, they need to be well-bracketed. If
+the adversary were able to store the return pointer from the first call and
+invoke it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on
+top of its stack and the assertion would fail. However, the security measures in
+Section~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
+the return pointer is local, so it can only be stored on the stack, but the part
+of the stack that is accessible to the adversary is cleared before the second
+invocation. In fact, the following lemma shows that there are also no other
+attacks that can break well-bracketedness of this example, i.e.\ the assertion
+never fails. It is similar to the two previous lemmas:
 
@@
 
-The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}}, is a faithful translation
-of a tricky example known from the
+The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}},
+is a faithful translation of a tricky example known from the
 literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
-parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}. \texttt{\footnotesize{g1}} is a closure generator that
-generates closures with one variable $x$ set to $0$ in its environment and
-\texttt{\footnotesize{f4}} as the program (note that we here use knowledge about the stack usage to make optimizations to the calling convention).
-\texttt{\footnotesize{f4}} expects one argument, a callback.
-The
-program sets $x$ to $0$ and calls the callback. When it returns, \texttt{\footnotesize{f4}}
-sets $x$ to $1$ and calls the callback a second time. When it returns from the
-callback, it asserts that $x$ is $1$ and returns. This example is more
-complicated than the previous ones because it involves a closure invoked by the
-adversary and an adversary callback invoked by us. As explained in
-Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
-to properly check the stack pointer received by our closure from the adversary
-(it must have the correct permission and fit with our calling convention) and (2)
-require that the adversary callback is global.
+parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
+\texttt{\footnotesize{g1}} is a closure generator that generates closures with
+one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
+the program (note that we can omit some calling convention security measures
+because the stack is not used or passed to the adversary in the closure
+generator). \texttt{\footnotesize{f4}} expects one argument, a callback. It sets
+$x$ to $0$ and calls the callback. When it returns, it sets $x$ to $1$ and calls
+the callback a second time. When it returns again, it asserts that $x$ is $1$
+and returns. This example is more complicated than the previous ones because it
+involves a closure invoked by the adversary and an adversary callback invoked by
+us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means we
+need to check (1) that the stack pointer that the closure receives from the
+adversary (it must have the correct permission and fit with our calling
+convention) and (2) that the adversary callback is global.
 
@@
 the adversary to transfer the return pointer from the point where it reinvokes
-the closure to where the closure reinvokes the callback, then the assertion can
-be made to fail. Similarly, if there were any way for the adversary to store a
-stack pointer or trick the trusted code into preserving it across an invocation,
-the assertion can likely be made to fail too. However, our calling convention
-guarantees that there is no way any of this can happen, as we prove in the
-following lemma.
+the closure to where the closure reinvokes the callback, then the assertion
+could be made to fail. Similarly, if there were any way for the adversary to
+store a stack pointer or trick the trusted code into preserving it across an
+invocation, the assertion can likely be made to fail too. However, our calling
+convention prevents any of this from happening, as we prove in the following
+lemma.
 

### 2017-02-16 csf-2017/paper.tex
Fix the sentence that Lau noticed was broken.
LR changes

@@
 For technical reasons, we do not actually remove a revoked temporary region from
-the world, but we turn it a special revoked region that exists for this purpose.
+the world, but we turn it into a special revoked region that exists for this purpose.
 Such a revoked region contains no state transition system and puts no
@@
 \begin{mathpar}
-  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
-    (s,s') \in \phi_\pub }
-  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
+  \inferrule{ (s,s') \in \phi_\pub }
+  {  (v,s',\phi_\pub,\phi,H) \futurewk (v,s,\phi_\pub,\phi,H) }
   \and
@@
 \begin{mathpar}
-  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
-              (s,s') \in \phi} 
-  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
+  \inferrule{ (s,s') \in \phi} 
+  {  (v,s',\phi_\pub,\phi,H) \futurestr (v,s,\phi_\pub,\phi,H) }
   \and
@@
 The logical relation now defines semantically when values, terms, and
-configurations are for execution by the adversary, i.e. without the risk that
-they will be able to use it to break protocols in the world. The definition of
-the logical relation is found in \figurename{}s~\ref{fig:logrel}
+configurations can safely be used or executed by the adversary, i.e. without the
+risk that they will be able to use it to break protocols in the world. The
+definition of the logical relation is found in \figurename{}s~\ref{fig:logrel}
 and~\ref{fig:perm-conds} and we provide some explanations in the following
@@
     \begin{array}{r@{}lr@{}l}
-    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
+    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
     c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
@@
 %\label{sec:conclusion}
-\todo[inline]{Insert the correct URL for the technical report in the bibtex file!}
 

### 2017-02-16 csf-2017/paper.tex
Pass over LR section done

@@
 The \emph{observation relation} $\observations$ defines configurations that we
-consider safe. Roughly speaking, a configuration is safe if, whenever it halts,
+consider safe. Roughly speaking, a configuration is safe if whenever it halts,
 there is a private future world that the resulting memory satisfies. Notice that
@@
 
-For a capability with write permission, $\writeCond{}$ must be satisfied for the
-capability's range of authority. Such a capability may be used to write any
-word, so the region governing the relevant memory must (at least) allow any safe
-word to be stored. More precisely, for permit-write-local capabilities it needs
-to allow arbitrary safe values to be stored, while for write capabilities,
-storing values that are both non-local and safe is sufficient. Like before, we
-use $\var{localityReg}$ to pick an appropriate region for the capability's
-locality. The condition for the region to be address-stratified is a technical
-requirement, defined in the technical appendix. In the value relation, the write
-condition is used with $\iota^\pwl$ for write-local capabilities. For write
-capabilities, it is used with another standard region
+For a capability with write permission, $\writeCond{}$ must be
+satisfied for the capability's range of authority. Such a capability
+may be used to write any word, so the region governing the relevant
+memory must (at least) allow any safe word to be stored. More
+precisely, for write-local capabilities it needs to allow arbitrary
+safe values to be stored, while for write capabilities, storing values
+that are both non-local and safe is sufficient.  This is done in the
+value relation by the regions the write condition is applied with. For
+write-local capabilities $\iota^\pwl$ is used, and for write
+capabilities we use another standard region
 $\iota^\nwl_{\start,\addrend}$ which requires that the words in
-$[\start,\addrend]$ are non-local and safe.
+$[\start,\addrend]$ are non-local and safe.  Like before, we use
+$\var{localityReg}$ to pick an appropriate region for the capability's
+locality. The condition for the region to be address-stratified is a
+technical requirement, defined in the technical appendix.
 
@@
 called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}. To
-understand this, consider that the theorem says that the capability
+understand this, consider that the theorem says the capability
 $((\perm,\gl),\start,\addrend,\addr)$ is safe as a program counter, without any

### 2017-02-16 csf-2017/paper.tex
Vertically flip the activation record in Figure 4.

@@
     \draw[draw=gray!80] (6,0.5) -- (6,4);
-    \draw (0,3.5) rectangle (6,3) node[pos=.5] {\footnotesize previous program counter};
-    \draw (0,3) rectangle (6,2.5) node[pos=.5] {\footnotesize previous stack pointer};
-    \draw (0,2.5) rectangle (6,1) node[pos=.5] {\footnotesize restore instructions};
+    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous program counter};
+    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous stack pointer};
+    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
     \draw (-3,2.25) node (rp) {return pointer};

### 2017-02-16 csf-2017/paper.tex
Adapt the scall description to the current state of Figure 4.

@@
 capability in register $r$ in the manner described above. That is, it pushes
-local state (the contents of registers $\overline{r_{\var{priv}}}$), stack
-pointer and return pointer to the stack, sets up the necessary on stack return
+local state (the contents of registers $\overline{r_{\var{priv}}}$), return
+pointer and stack pointer to the stack, sets up the necessary on stack return
 code and creates an $\entry$ return pointer, restricts the stack pointer, clears

### 2017-02-16 csf-2017/paper.tex
Small textual change.

@@
 \]
-  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1} 
-  if
+  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
+  Then if
   \[
@@
 
-As explained in Section~\ref{sec:stack-and-return-pointer}, the callback must be
-global, essentially to make sure that it is not allocated on the stack where it
-might contain the adversary's original stack pointer. If not, it could break the
+As explained in Section~\ref{sec:stack-and-return-pointer}, the
+macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
+is global, essentially to make sure it is not allocated on the stack where it
+might contain old stack pointers or return pointers. If not, it could break the
 encapsulation of our local stack frame. In the proof of

### 2017-02-16 csf-2017/paper.tex
figure 3 and scall

@@
     \draw[draw=gray!80] (6,0.5) -- (6,4);
-    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous program counter};
-    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous stack pointer};
+    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous stack pointer};
+    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous program counter};
     \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
-    \draw (-3,2.25) node (rp) {return pointer};
-    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,2.25);
+    \draw (-3,3.25) node (rp) {return pointer};
+    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,3.25);
   \end{tikzpicture}
@@
 \end{description}
-\textbf{Reusable macro instructions} We define a number of reusable macros
-capturing the calling convention and other conveniences. All macros that uses
-the stack assumes a stack pointer in register $r_\stk$.
-\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
-$\var{name}$ from the linking table and stores it in register $r$.
-\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
-would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
-invoked by the adversary and prepares the received stack pointer by checking
-that it has permission $\rwlx$.
-\texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
-capability in register $r$ in the manner described above. That is, it pushes
-local state (the contents of registers $\overline{r_{\var{priv}}}$), return
-pointer and stack pointer to the stack, sets up the necessary on stack return
-code and creates an $\entry$ return pointer, restricts the stack pointer, clears
-the unused part of the stack, clears the necessary registers and jumps to $r$.
-Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
-clears all the memory the capability in register $r$ has authority over.
-\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
-$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
-register $r$ is a global capability. \texttt{\footnotesize{crtcls
-    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
-closure's code and a new environment is allocated (using malloc) where the
-contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
+\textbf{Reusable macro instructions} We define a number of reusable
+macros capturing the calling convention and other conveniences. All
+macros that uses the stack assumes a stack pointer in register
+$r_\stk$.  \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
+capability related to $\var{name}$ from the linking table and stores
+it in register $r$.  \texttt{\footnotesize{push $r$}} and
+\texttt{\footnotesize{pop $r$}} work as one would
+expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
+invoked by the adversary and prepares the received stack pointer by
+checking that it has permission $\rwlx$.  \texttt{\footnotesize{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}}
+jumps to the capability in register $r$ in the manner described
+above. That is, it pushes local state (the contents of registers
+$\overline{r_{\var{priv}}}$), the necessary on stack return code,
+return pointer, stack pointer to the stack, and creates an $\entry$
+return pointer, restricts the stack pointer, clears the unused part of
+the stack, clears the necessary registers and jumps to $r$.  Upon
+return, the private state is restored. \texttt{\footnotesize{mclear
+    $r$}} clears all the memory the capability in register $r$ has
+authority over.  \texttt{\footnotesize{rclear $\var{regSet}$}} clears
+all the registers in $\var{regSet}$. \texttt{\footnotesize{reqglob
+    $r$}} checks whether the word in register $r$ is a global
+capability. \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}}
+allocates a closure where $r$ points to the closure's code and a new
+environment is allocated (using malloc) where the contents of
+$\overline{r_i}$ is stored. In the code referred to by $r$, an
 implicit fetch happens when an instruction refers to $x_i$.

### 2017-02-16 csf-2017/paper.tex
style corrections

@@
 \begin{itemize}
-\item Formalise a simple but representative capability machine featuring local
+\item Formalize a simple but representative capability machine featuring local
   capabilities, and its operational semantics
@@
 \label{sec:examples}
-In this section, we demonstrate how our formalisation of capability safety
+In this section, we demonstrate how our formalization of capability safety
 allows us to prove local-state encapsulation and control-flow correctness
@@
 
-\item Some compartmentalisation result?
+\item Some compartmentalization result?
 \end{itemize}
@@
 is not made very explicit. As future work, we consider looking at a more
-explicit and useful way to formalise well-bracketed control flow. The idea would
+explicit and useful way to formalize well-bracketed control flow. The idea would
 be to define a variant of our capability machine with built-in call and return
@@
 We
-generalise this idea to reason about local capabilities: validity of local
+generalize this idea to reason about local capabilities: validity of local
 capabilities is allowed to be monotone with respect to a weaker future-world
@@
 
-In this section, we summarise how our work relates to previous work. We do not
+In this section, we summarize how our work relates to previous work. We do not
 repeat the work already discussed in Section~\ref{sec:discussion}.

### 2017-02-16 csf-2017/paper.tex
pass over example section

@@
 From another point of view, the Fundamental Theorem also serves as a sanity
-check of the logical relation as it guarantees that a large class of
+check of the logical relation as it guarantees a large class of
 capabilities actually inhabits it.
@@
 \texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
-\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's
-encapsulation of local state. They are very similar: both store some local
-state, call an untrusted piece of code ($\adv$), and then test whether the local
-state is unchanged. They differ in the way they do this. Program
-\texttt{\footnotesize{f1}} assumes access to a stack pointer and uses the stack
-to store its local state and the stack-based calling convention
-\texttt{\footnotesize{scall}} to call the adversary. On the other hand,
-\texttt{\footnotesize{f2}} uses malloc to allocate memory for its local state
-and uses an activation-record based calling convention (described in the
-technical appendix) to run the adversarial code.
+\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability
+machine's encapsulation of local state. They are very similar: both
+store some local state, call an untrusted piece of code ($\adv$), and
+then test whether the local state is unchanged. They differ in the way
+they do this. Program \texttt{\footnotesize{f1}} uses our stack-based
+calling convention (captured by \texttt{\footnotesize{scall}}) to call
+the adversary, so it can use the available stack to store its local
+state.  On the other hand, \texttt{\footnotesize{f2}} uses malloc to
+allocate memory for its local state and uses an activation-record
+based calling convention (described in the technical appendix) to run
+the adversarial code.
 
@@
 \texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
-fails. To this end, we assume they have access to a flag (an address in memory)
-that they set to $1$ if the assertion fails and then halt. The correctness lemma
+fails. To this end, we assume they have access to a flag (an address in memory).
+If the assertion fails, then the flag is set to $1$ and execution halts. The correctness lemma
 for \texttt{\footnotesize{f1}} then states:
@@
 one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
-the program (note that we can omit some calling convention security measures
+the program (note we can omit some calling convention security measures
 because the stack is not used or passed to the adversary in the closure
@@
 $x$ to $0$ and calls the callback. When it returns, it sets $x$ to $1$ and calls
-the callback a second time. When it returns again, it asserts that $x$ is $1$
+the callback a second time. When it returns again, it asserts $x$ is $1$
 and returns. This example is more complicated than the previous ones because it
@@
 us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means we
-need to check (1) that the stack pointer that the closure receives from the
+need to check (1) the stack pointer that the closure receives from the
 adversary (it must have the correct permission and fit with our calling
-convention) and (2) that the adversary callback is global.
+convention) and (2) the adversary callback is global.
 
@@
 world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
-adversary is again allowed to rely on its old local state which means that we
+adversary is again allowed to rely on its old local state which means we
 need to guarantee that the local state is unchanged. This is important because
@@
 and the adversary is allowed to allocate the activation record on the stack
-(just like we do) so that they can store and recover his old stack pointer after
+(just like we do) so they can store and recover their old stack pointer after
 \texttt{\footnotesize{f4}} returns.

### 2017-02-16 csf-2017/paper.tex
Shorten, rephrase.

--- 2017-02-16 csf-2017/paper.tex
@@
     \draw[draw=gray!80] (6,0.5) -- (6,4);
-    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous stack pointer};
-    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous program counter};
+    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous program counter};
+    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous stack pointer};
     \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
-    \draw (-3,3.25) node (rp) {return pointer};
-    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,3.25);
+    \draw (-3,2.25) node (rp) {return pointer};
+    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,2.25);
   \end{tikzpicture}
@@
 \end{description}
-\textbf{Reusable macro instructions} We define a number of reusable
-macros capturing the calling convention and other conveniences. All
-macros that uses the stack assumes a stack pointer in register
-$r_\stk$.  \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
-capability related to $\var{name}$ from the linking table and stores
-it in register $r$.  \texttt{\footnotesize{push $r$}} and
-\texttt{\footnotesize{pop $r$}} work as one would
-expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
-invoked by the adversary and prepares the received stack pointer by
-checking that it has permission $\rwlx$.  \texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}}
-jumps to the capability in register $r$ in the manner described
-above. That is, it pushes local state (the contents of registers
-$\overline{r_{\var{priv}}}$), the necessary on stack return code,
-return pointer, stack pointer to the stack, and creates an $\entry$
-return pointer, restricts the stack pointer, clears the unused part of
-the stack, clears the necessary registers and jumps to $r$.  Upon
-return, the private state is restored. \texttt{\footnotesize{mclear
-    $r$}} clears all the memory the capability in register $r$ has
-authority over.  \texttt{\footnotesize{rclear $\var{regSet}$}} clears
-all the registers in $\var{regSet}$. \texttt{\footnotesize{reqglob
-    $r$}} checks whether the word in register $r$ is a global
-capability. \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}}
-allocates a closure where $r$ points to the closure's code and a new
-environment is allocated (using malloc) where the contents of
-$\overline{r_i}$ is stored. In the code referred to by $r$, an
+\textbf{Reusable macro instructions} We define a number of reusable macros
+capturing the calling convention and other conveniences. All macros that uses
+the stack assumes a stack pointer in register $r_\stk$.
+\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
+$\var{name}$ from the linking table and stores it in register $r$.
+\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} work as one
+would expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
+invoked by the adversary and prepares the received stack pointer by checking
+that it has permission $\rwlx$.
+\texttt{\footnotesize{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
+capability in register $r$ in the manner described above. That is, it pushes
+local state (the contents of registers $\overline{r_{\var{priv}}}$), return
+pointer and stack pointer to the stack, sets up the necessary on stack return
+code and creates an $\entry$ return pointer, restricts the stack pointer, clears
+the unused part of the stack, clears the necessary registers and jumps to $r$.
+Upon return, the private state is restored. \texttt{\footnotesize{mclear $r$}}
+clears all the memory the capability in register $r$ has authority over.
+\texttt{\footnotesize{rclear $\var{regSet}$}} clears all the registers in
+$\var{regSet}$. \texttt{\footnotesize{reqglob $r$}} checks whether the word in
+register $r$ is a global capability. \texttt{\footnotesize{crtcls
+    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
+closure's code and a new environment is allocated (using malloc) where the
+contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
 implicit fetch happens when an instruction refers to $x_i$.
@@
 is global, essentially to make sure it is not allocated on the stack where it
-might contain old stack pointers or return pointers. If not, it could break the
-encapsulation of our local stack frame. In the proof of
+might contain old stack pointers or return pointers. Otherwise, the
+encapsulation of our local stack frame could be broken. In the proof of
 Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
 callback in a world that is only a private future world of the one where we
-received the callback, precisely because we have invalidated the local state of
-the adversary (particularly its previous stack and return capabilities). The
-callback is still valid in this private future world, but only because we know
-that it is global.
+received the callback, precisely because we have invalidated the adversary's
+local state (particularly their old stack and return capabilities). The callback
+is still valid in this private future world, but only because we know that it is
+global.
 
@@
 compared to the previous lemmas. In this lemma, the adversary assumes control
-first with a capability for the closure creator \texttt{\footnotesize{g1}}. This means we need
-to check that all arguments are safe to use and that we clean up before
-returning in the end. The inversion of control poses an interesting challenge
-when it comes to reasoning about the adversary's local state during the
-execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
-the adversary should not be able to rely on its local state from before the call
-of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
-world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
-adversary is again allowed to rely on its old local state which means that we
-need to guarantee that the local state is unchanged. This is important because
-the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
-and the adversary is allowed to allocate the activation record on the stack
-(just like we do) so that they can store and recover his old stack pointer after
-\texttt{\footnotesize{f4}} returns.
-
-In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in the
-proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
-all the situations where control is passed to the adversary. The potentially
-local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
-$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
-constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
-by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
-we pass away and static temporary regions (a region that accepts exactly one
-memory) for our local stack as well as the adversary's local stack. The given
-worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
-to return to us. None of these worlds can have masked the temporary regions of
-$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
-regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
-$W_6$ a public future world of $W_1$. It is therefore safe to use the return
-pointer to return to the adversary, since we have restored validity of any local
-state they might have stashed away.
+first with a capability for the closure creator \texttt{\footnotesize{g1}}. This
+means we need to check that all arguments are safe to use and that we clean up
+before returning in the end. The inversion of control makes it challenging to
+reason about the adversary's local state during the execution of
+\texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}}
+and the callbacks, the adversary cannot use their local state from before the
+call of \texttt{\footnotesize{f4}}. This is modeled by revoking the temporary
+regions of the world we get at the start of \texttt{\footnotesize{f4}}. However,
+when \texttt{\footnotesize{f4}} returns, the adversary's old local state becomes
+available again, since they are allowed to allocate their activation record on
+the stack (just like we do) so that they can store and recover their old stack
+pointer after \texttt{\footnotesize{f4}} returns. This means that we need to
+restore validity of the adversary's old local state, which we achieve
+essentially by clearing the stack.
+
+In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
+the proof of Lemma~\ref{lem:correctness-g1} by constructing suitable worlds for
+all situations where we pass control to the adversary. The potentially local
+return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe
+in $W_1$, so it can only be used in public future worlds. Let us see how $W_6$
+is constructed as a private future world of $W_5$ but also a public future world
+of $W_1$. The worlds $W_2$ and $W_4$ are constructed by revoking all temporary
+regions and adding a $\iota^\pwl$ region for the stack we pass away and static
+temporary regions (which accept exactly one memory) for our local stack as well
+as the adversary's old local stack. The given worlds $W_3$ and $W_5$ are public
+future worlds in which the adversary chooses to return to us. None of these
+worlds can have replaced the temporary regions of $W_1$ with permanent regions.
+Therefore, by revoking the temporary regions of $W_5$ and reinstating the
+temporary regions of $W_1$, we can construct $W_6$ as a private future world of
+$W_5$ but also a public future world of $W_1$. Because of the latter, it is safe
+to use the local return pointer to return to the adversary, since we have
+restored validity of any local state they might have stashed away.
 
@@
 \end{tikzpicture}
-  \caption{Illustration of some of the worlds in the proof of
-    Lemma~\ref{lem:correctness-g1}. The first row of worlds are worlds
-  we construct during the proof. The second row of worlds are worlds
-  we are given during the proof. $W_6$ is constructed such that $W_6
-  \futurewk W_1$.}
+\caption{Illustration of the worlds in the proof of
+  Lemma~\ref{lem:correctness-g1}. In the proof, the above row of worlds are
+  constructed by us, while the below worlds are given. $W_6$ is constructed such
+  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
   \label{fig:worlds-in-pf-g1}
@@
 \noindent\textbf{Calling convention}\\
-\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow
-correctness, we do not prove a general theorem that shows this, because it is not clear what such a theorem should state. Formulations in terms of a control-flow graph, like the one by
+\emph{Formulating control flow correctness} While we claim that our calling
+convention enforces control-flow correctness, we do not prove a general theorem
+that shows this, because it is not clear what such a theorem should look like.
+Formulations in terms of a control-flow graph, like the one by
 \cite{abadi_control-flow_2005}, do not take into account temporal properties,
 like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
-In fact, our examples show that our logical relation imply a form of
-control-flow correctness that is stronger than such formulations, although this
-is not made very explicit. As future work, we consider looking at a more
-explicit and useful way to formalise well-bracketed control flow. The idea would
-be to define a variant of our capability machine with built-in call and return
-instructions and well-bracketed control flow built-in to the operational
-semantics, and then prove that compiling such code to our machine using our
-calling convention is fully abstract~\cite{abadi_protection_1998}.
-
-\emph{Performance and the requirement for stack clearing} 
-The additional security measures of the calling convention described in
-Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to
-a calling convention that does not provide security. However, the security measures generally consist of a
-few checks or register clearings on every boundary crossing between trusted code
-and adversary, which should impose an acceptable performance overhead. The only
-exception is the requirements for stack clearing which we occur in two
-situations: when returning to the adversary and when invoking an adversary
-callback. As we have explained, it is not sufficient to clear just the part of
-the stack that we have used ourselves, but we need to clear all of the stack that we are not using ourselves. 
-In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be cleared.
-We believe this is a common requirement for typical usage scenarios of local capabilities and
-capability machines like CHERI should consider to provide special support for
-this requirement, in the form of a highly-optimized instruction for erasing
-a large block of memory.
-
-\emph{Modularity} It is important that our calling convention is modular.
-This means that the requirements we have on callbacks, return
-pointers etc. provided to trusted code by an adversary are also satisfied by
-callbacks, return pointers etc. that we construct ourselves. For example, our
-return pointers are local capabilities because they must point to
-memory where we can store the old stack pointer, but the adversary's return
-pointers are also allowed to be local. Adversary callbacks on the other hand are
-required to be global but the callbacks we construct ourselves are also
-allocated on the heap and global. In other words, we do not assume that our code
-has some sort of special status and it is possible for the adversary to protect
-themselves against us the same way we protect against them.
-\\\\
+In fact, our examples show that our logical relation imply a stronger form of
+control-flow correctness than such formulations, although this is not made very
+explicit. As future work, we consider looking at a more explicit and useful way
+to formalise control flow correctness. The idea would be to define a variant
+of our capability machine with call and return instructions and
+well-bracketed control flow built-in to the operational semantics, and then
+prove that compiling such code to our machine using our calling convention is
+fully abstract~\citep{abadi_protection_1998}.
+
+\emph{Performance and the requirement for stack clearing} The additional
+security measures of the calling convention described in
+Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
+calling convention without security guarantees. However, most security measures
+require only a few atomic checks or register clearings on boundary crossings
+between trusted code and adversary, which should produce an acceptable
+performance overhead. The only exception are the requirements for stack clearing
+that we have in two situations: when returning to the adversary and when
+invoking an adversary callback. As we have explained, we need to clear all of
+the stack that we are not using ourselves, not just the part that we have
+actually used. In other words, on every boundary cross between trusted code and
+adversary code, a potentially large region of memory must be cleared. We believe
+this is actually a common requirement for typical usage scenarios of local
+capabilities and capability machines like CHERI should consider to provide
+special support for this requirement, in the form of a highly-optimized
+instruction for erasing a large block of memory.
+
+\emph{Modularity} It is important that our calling convention is modular, i.e.
+we do not assume that our code is specially privileged w.r.t. the adversary, and
+they can apply the same measures to protect themselves from us as we do to
+protect ourselves from them. More concretely, the requirements we have on
+callbacks and return pointers received from the adversary are also satisfied by
+callbacks and return pointers that we pass to them. For example, our return
+pointers are local capabilities because they must point to memory where we can
+store the old stack pointer, but the adversary's return pointers are also
+allowed to be local too. Adversary callbacks on the other hand are required to
+be global but the callbacks we construct ourselves are allocated on the heap and
+also global. \\\\
 \noindent\textbf{Logical relation}\\
@@
 safe observations when plugging safe values and define safe terms as the set of
-terms that can be plugged into safe evaluation contexts to produce safe values.
-This is an alternative to more direct definitions where safe terms are defined
-as terms that evaluate to safe values. An advantage of biorthogonality is that
-it scales better to languages with control effects like call/cc. Our definitions
-can be seen as a variant of biorthogonality, where we take only a single
-orthogonal closure: we do not define safe evaluation contexts but immediately
-define safe terms as those that produce safe observations when plugged with safe
-values. This is natural because we model arbitrary assembly code that does not
-necessarily respect a particular calling convention. As a result, return
-pointers are in principle values like all others and there is no reason to treat
-them specially in the logical relation.
-
-It is interesting to note that \citet{Hur:2011:KLR:1926385.1926402} also build a
-step-indexed, Kripke, biorthogonal logical relation for an assembly language
-(for reasoning about correct compilation from ML to assembly), but because they
-only model non-adversarial code that treats return pointers specially according
-to a particular calling convention, they can use standard biorthogonality rather
-than a single orthogonal closure like us.
+terms that can be plugged into safe evaluation contexts to produce safe
+observations. This is an alternative to more direct definitions where safe terms
+are defined as terms that evaluate to safe values. An advantage of
+biorthogonality is that it scales better to languages with control effects like
+call/cc. Our definitions can be seen as a variant of biorthogonality, where we
+take only a single orthogonal closure: we do not define safe evaluation contexts
+but immediately define safe terms as those that produce safe observations when
+plugged with safe values. This is natural because we model arbitrary assembly
+code that does not necessarily respect a particular calling convention. As a
+result, return pointers are in principle values like all others and there is no
+reason to treat them specially in the logical relation.
+
+Interestingly, \citet{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
+Kripke logical relation for an assembly language (for reasoning about correct
+compilation from ML to assembly), but because they only model non-adversarial
+code that treats return pointers according to a particular calling convention,
+they can use standard biorthogonality rather than a single orthogonal closure
+like us.
 
 \emph{Public/private future worlds} A novel aspect of our logical relation is
-how we model the temporary nature of local capabilities using public/private
+how we model the special nature of local capabilities using public/private
 future worlds. The main insight is that the temporary, revokable nature of these
@@
 unstorable status of evaluation contexts in lambda calculi without control
-effects (of which well-bracketed control flow is a consequence). To reason about
-code that relies on this (particularly, the original awkward example),
-\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
-contexts using Kripke worlds with public and private future world relations.
-Essentially, they allow relatedness of evaluation contexts to be monotone with
-respect to a weaker future world relation (public) than relatedness of values,
-formalising the idea that it is safe to make temporary internal state
-modifications (private world transitions, that invalidate the continuation,
-but not other values) while an expression is performing internal steps, as long
-as the code returns to a stable state (i.e. transitions to a public future world
-of the original) before returning (i.e.
-invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation).
-Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
-We
-generalise this idea to reason about local capabilities: validity of local
-capabilities is allowed to be monotone with respect to a weaker future-world
-relation than other values, which we can exploit to distinguish between state
-changes that are always safe (public future worlds) and changes that are only
-valid if we clear all local capabilities (private future worlds). Our future
-world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
-the awkward example uses exactly the same state transition system),
-but they turn up in an entirely different place in the logical relation: rather
-than using public future worlds for the special syntactic category of evaluation
-contexts, they are used in the value relation depending on the
-locality of the capability at hand. Additionally, our worlds are a bit more
-complex because, to allow local memory capabilities and write-local
-capabilities, they can contain (revokable) temporary regions that are only
-monotonous w.r.t. public future worlds, while DNB's worlds are entirely
+effects. To reason about code that relies on this (particularly, the original
+awkward example), \citet{Dreyer:jfp12} (DNB) formally capture the special status
+of evaluation contexts using Kripke worlds with public and private future world
+relations. Essentially, they allow relatedness of evaluation contexts to be
+monotone with respect to a weaker future world relation (public) than
+relatedness of values. This formalises the idea that it is safe to make
+temporary internal state modifications (private world transitions, that
+invalidate the continuation, but not other values) while a term is performing
+internal steps, as long as it returns to a stable state (i.e. transitions to a
+public future world of the original) before returning. We generalise this idea
+to reason about local capabilities: validity of local capabilities is allowed to
+be monotone with respect to a weaker future-world relation than other values,
+which we can exploit to distinguish between state changes that are always valid
+(public future worlds) and changes that are only valid if we clear local
+capabilities (private future worlds). Our future world relations are similar to
+\citet{Dreyer:jfp12}'s (for example, our proof of the awkward example uses the
+exact same state transition system), but they turn up in an entirely different
+place in the logical relation: rather than using public future worlds for the
+special syntactic category of evaluation contexts, they are used in the value
+relation depending on the locality of the capability at hand. Additionally, our
+worlds are more complex because, to allow local memory capabilities and
+write-local capabilities, they can contain (revokable) temporary regions that
+are only monotonous w.r.t. public future worlds, while DNB's worlds are entirely
 permanent.
 
-\emph{Local capabilities in high-level languages} It is worth mentioning that
+\emph{Local capabilities in high-level languages} We point out that
 local capabilities are quite similar to a feature proposed for the high-level
@@
 later use. We believe reasoning about such values will require techniques similar 
-to the ones we have used for local capabilities.
+to what we provide for local capabilities.
 
@@
 
-In this section, we summarise how our work relates to previous work. We do not
-repeat the work already discussed in Section~\ref{sec:discussion}.
-
-Capability machines originate with Dennis and Van
-Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
+Finally, we summarise how our work relates to previous work. We do not
+repeat the work we discussed in Section~\ref{sec:discussion}.
+
+Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to
 \citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
@@
 model, modeled mainly after the
-M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter pointers
+M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers
 resemble the M-Machine's) and
-CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
+CHERI~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
 memory and local capabilities resemble CHERI's). The latter is a
@@
 backwards compatibility and gradual adoption. As discussed, our local
-capabilities differ from CHERI's by being able to cross module boundaries, 
+capabilities can cross module boundaries, 
 contrary to what is enforced by CHERI's default CCall implementation.
 
-There are plenty of other papers\lau{If there are plenty, then I guess we should cite more than one?}
- that enforce well-bracketed control flow at a
-low level, but most are restricted to preventing particular types of attacks and
-enforce only partial correctness of control flow. This includes particularly the
-line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
-This work uses a quite different attacker model than us: they assume an attacker
-that is not able to execute code, but can overwrite arbitrary data at any time
-during execution (as a model for buffer overflows). By checking the address of
-every indirect jump and using memory access control to prevent overwriting code,
-this work enforces what they call control-flow integrity, formalized as the
-property that every jump will follow a legal path in the control-flow graph.
-As discussed in Section~\ref{sec:discussion}, this integrity property does not
-take into account temporal properties and seems hard to use for reasoning.
+Plenty of other papers\lau{If there are plenty, then I guess we should cite more
+  than one?} enforce well-bracketed control flow at a low level, but most are
+restricted to preventing particular types of attacks and enforce only partial
+correctness of control flow. This includes particularly the line of work on
+\emph{control-flow integrity}~\citep{abadi_control-flow_2005}. Those use a quite
+different attacker model than us: they assume an attacker that is not able to
+execute code, but can overwrite arbitrary data at any time during execution (to
+model buffer overflows). By checking the address of every indirect jump and
+using memory access control to prevent overwriting code, this work enforces what
+they call control-flow integrity, formalized as the property that every jump
+will follow a legal path in the control-flow graph. As discussed in
+Section~\ref{sec:discussion}, such a property ignores temporal properties and
+seems hard to use for reasoning.
 
@@
 with a primitive notion of compartments) and we do not use a trusted stack
-manager component, but a decentralized calling convention based on local
+manager, but a decentralized calling convention based on local
 capabilities. Also, both prove a secure compilation result from a high-level
 language, which clearly implies a general form of control-flow correctness,
-while we define a sound logical relation that can be used to reason about
+while we define a logical relation that can be used to reason about
 specific programs that rely on well-bracketed control flow.
@@
 capability safety in a high-level JavaScript-like lambda calculus. Our
-Fundamental Theorem is similar to theirs and can be seen as a formulation of
+Fundamental Theorem is similar to theirs and expresses
 capability safety of the capability machine. Because we are not interested in
@@
 result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
-correctness). He also proves soundness of control-flow integrity (see above)
-adapted to the capability machine, a result that seems orthogonal to the fact
-that the machine uses capabilities.
+correctness). He also adapts control-flow integrity (see above) to the machine
+and shows soundness, seemingly without relying on capabilities.
 

### 2017-02-16 csf-2017/paper.tex
Oops, fix the merge.

@@
 % programs which means that the continuation is always invoked in the 
-% same way. Our realisation was that anything we pass as an argument 
+% same way. Our realization was that anything we pass as an argument 
 % in a "call" can be used as a continuation, so the continuation 
@@
 % \item the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
-% \item some compartmentalisation result?
+% \item some compartmentalization result?
 % \item whatever else we do..
@@
 With the logical relation defined, we can now state the fundamental theorem of
-our logical relation: a strong theorem that formalises the guarantees offered by
+our logical relation: a strong theorem that formalizes the guarantees offered by
 the capability machine. Essentially, it says that a capability that only grants
@@
 explicit. As future work, we consider looking at a more explicit and useful way
-to formalise control flow correctness. The idea would be to define a variant
+to formalize control flow correctness. The idea would be to define a variant
 of our capability machine with call and return instructions and
@@
 Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
-calling convention without security guarantees. However, most security measures
+calling convention without security guarantees. However, most of our security measures
 require only a few atomic checks or register clearings on boundary crossings
@@
 monotone with respect to a weaker future world relation (public) than
-relatedness of values. This formalises the idea that it is safe to make
+relatedness of values. This formalizes the idea that it is safe to make
 temporary internal state modifications (private world transitions, that
@@
 internal steps, as long as it returns to a stable state (i.e. transitions to a
-public future world of the original) before returning. We generalise this idea
+public future world of the original) before returning. We generalize this idea
 to reason about local capabilities: validity of local capabilities is allowed to
@@
 
-Finally, we summarise how our work relates to previous work. We do not
+Finally, we summarize how our work relates to previous work. We do not
 repeat the work we discussed in Section~\ref{sec:discussion}.

### 2017-02-16 csf-2017/paper.tex
'that' related changes

--- 2017-02-16 csf-2017/paper.tex
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine that the adversary stores his stack pointer, invokes trusted
+pointers. Imagine the adversary stores his stack pointer, invokes trusted
 code that uses part of the stack to store private data and then invokes the
@@
 With the logical relation defined, we can now state the fundamental theorem of
-our logical relation: a strong theorem that formalises the guarantees offered by
-the capability machine. Essentially, it says that a capability that only grants
+our logical relation: a strong theorem that formalizes the guarantees offered by
+the capability machine. Essentially, it says a capability that only grants
 safe authority is capability safe as a program counter.
@@
 involves a closure invoked by the adversary and an adversary callback invoked by
-us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means we
-need to check (1) the stack pointer that the closure receives from the
-adversary (it must have the correct permission and fit with our calling
-convention) and (2) the adversary callback is global.
+us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means that we
+need to check (1) that the stack pointer that the closure receives from the
+adversary has write-local permission and (2) that the adversary callback is global.
 
@@
 compared to the previous lemmas. In this lemma, the adversary assumes control
-first with a capability for the closure creator \texttt{\footnotesize{g1}}. This means we need
+first with a capability for the closure creator \texttt{\footnotesize{g1}}. Consequently, we need
 to check that all arguments are safe to use and that we clean up before
@@
 world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
-adversary is again allowed to rely on its old local state which means we
+adversary is again allowed to rely on its old local state so we
 need to guarantee that the local state is unchanged. This is important because
@@
 formalising the idea that it is safe to make temporary internal state
-modifications (private world transitions, that invalidate the continuation,
+modifications (private world transitions, which invalidate the continuation,
 but not other values) while an expression is performing internal steps, as long

### 2017-02-16 csf-2017/paper.tex
Small rephrasings and shortenings.

--- 2017-02-16 csf-2017/paper.tex
@@
 and local-state encapsulation are either limited to only certain common forms of
-attacks (leading to an attack-defense arms race) or rely on techniques like
+attacks (leading to an attack-defense arms race) and/or rely on techniques like
 machine code rewriting
@@
 
-\emph{Capability machines} are a type of processors, intended to
-remediate these limitations by presenting a better security model at
+\emph{Capability machines} are a type of processors that
+remediate these limitations with a better security model at
 the hardware level. They are based on old ideas
@@
 machines tag every word (in the register file and in memory) to
-enforce a strict separation between numbers and capabilities, a kind
-of pointers carrying authority. Memory capabilities carry
+enforce a strict separation between numbers and capabilities (a kind
+of pointers that carry authority). Memory capabilities carry
 the authority to read and/or write to a range of memory
-locations. Moreover, all capability machines
-offer some form of \emph{object capabilities}, which represent the
+locations. There is also a form of \emph{object capabilities}, which represent the
 authority to invoke a piece of code without exposing the code's
@@
 
-In contrast to the situation for commodity processors, the security
-primitives on capability machines lend themselves well to the
-enforcement of properties like local-state encapsulation and
-control-flow correctness. Potentially, they will enable new compilation
-schemes that enforce such properties in a way that is efficient but
-also 100\% watertight (ideally evidenced by a mathematical proof,
-guaranteeing that we do not end up in a new attack-defense arms
-race). However, there is still a lot that needs to happen to attain
-this goal. For example, it is quite non-trivial to devise a
-compilation scheme adapted to the details of a specific language's
-notion of encapsulation (e.g., encapsulation of private member
-variables in OO languages often behaves quite differently than
-encapsulation of private state in ML-like languages). And even if such
-a scheme were defined, a formal proof depends on a formalization of
-the encapsulation provided by the capability machine at hand.
+Unlike commodity processors, capability machines lend themselves well to
+enforcing local-state encapsulation. Potentially, they will enable compilation
+schemes that enforce this property in an efficient but also 100\% watertight way
+(ideally evidenced by a mathematical proof, guaranteeing that we do not end up
+in a new attack-defense arms race). However, a lot needs to happen before we get
+there. For example, it is far from trivial to devise a compilation scheme
+adapted to the details of a specific source language's notion of encapsulation
+(e.g., private member variables in OO languages often behave quite differently
+than private state in ML-like languages). And even if such a scheme were
+defined, a formal proof depends on a formalization of the encapsulation provided
+by the capability machine at hand.
 
@@
 machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
-It drops the standard contiguous C stack and splits it into a central, trusted
-shadow stack, managed by trusted call and return instructions, and disjoint,
-per-component stacks. To prevent illegal use of stack references, the approach
-relies on \emph{local capabilities}, a type of capabilities offered by CHERI
-to \emph{temporarily} relinquish authority, namely for the duration of a function invocation whereafter the capability can be revoked.
-However, a lot of details
-remain to be investigated (e.g., how to combine a trusted shadow stack with
-(potentially) cross-domain function pointers in a scalable way?) and there is no
-argument that this enforcement of control-flow correctness is watertight.
+the standard contiguous C stack is split into a central, trusted stack, managed
+by trusted call and return instructions, and disjoint, private, per-compartment
+stacks. To prevent illegal use of stack references, the approach relies on
+\emph{local capabilities}, a type of capabilities offered by CHERI to
+\emph{temporarily} relinquish authority, namely for the duration of a function
+invocation whereafter the capability can be revoked. However, details are scarce
+and a lot remains to be investigated (e.g., how to combine disjoint stacks with
+cross-domain function pointers in a scalable way?) and there is no argument that
+the approach is watertight.
 
@@
 \item Define a novel calling convention enforcing control-flow correctness and
-  encapsulation of local state on the stack
-  (Section~\ref{sec:stack-and-return-pointer}). It relies solely on local
-  capabilities and does not require OS support (like a trusted stack or
-  call/return instructions). It also does not use separate per-component stacks
-  but a single, fairly standard, contiguous stack. It supports higher-order
-  cross-component calls (e.g., (potentially) cross-component function pointers)
-  and can be efficient assuming only one additional piece of processor support:
-  an efficient instruction for clearing a memory range.
+  encapsulation of stack frames (Section~\ref{sec:stack-and-return-pointer}). It
+  relies solely on local capabilities and does not require OS support (like a
+  trusted stack or call/return instructions). It also does not use separate
+  per-component stacks but a single, fairly standard, contiguous stack. It
+  supports higher-order cross-component calls (e.g., cross-component function
+  pointers) and can be efficient assuming only one additional piece of processor
+  support: an efficient instruction for clearing a range of memory.
 \item Prove a theorem that is the most general and powerful formulation of the
@@
   guarantees in a lambda calculus. The logical relation and capability safety
-  theorem is presented in Section~\ref{sec:logical-relation}.
+  theorem are presented in Section~\ref{sec:logical-relation}.
 \item Demonstrate our results by applying them to challenging examples,

### 2017-02-16 csf-2017/paper.tex

@@
 our logical relation: a strong theorem that formalizes the guarantees offered by
-the capability machine. Essentially, it says that a capability that only grants
+the capability machine. Essentially, it says a capability that only grants
 safe authority is capability safe as a program counter.
@@
 compared to the previous lemmas. In this lemma, the adversary assumes control
-first with a capability for the closure creator \texttt{\footnotesize{g1}}. This
-means we need to check that all arguments are safe to use and that we clean up
-before returning in the end. The inversion of control makes it challenging to
-reason about the adversary's local state during the execution of
-\texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}}
-and the callbacks, the adversary cannot use their local state from before the
-call of \texttt{\footnotesize{f4}}. This is modeled by revoking the temporary
-regions of the world we get at the start of \texttt{\footnotesize{f4}}. However,
-when \texttt{\footnotesize{f4}} returns, the adversary's old local state becomes
-available again, since they are allowed to allocate their activation record on
-the stack (just like we do) so that they can store and recover their old stack
-pointer after \texttt{\footnotesize{f4}} returns. This means that we need to
-restore validity of the adversary's old local state, which we achieve
-essentially by clearing the stack.
-
-In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
-the proof of Lemma~\ref{lem:correctness-g1} by constructing suitable worlds for
-all situations where we pass control to the adversary. The potentially local
-return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe
-in $W_1$, so it can only be used in public future worlds. Let us see how $W_6$
-is constructed as a private future world of $W_5$ but also a public future world
-of $W_1$. The worlds $W_2$ and $W_4$ are constructed by revoking all temporary
-regions and adding a $\iota^\pwl$ region for the stack we pass away and static
-temporary regions (which accept exactly one memory) for our local stack as well
-as the adversary's old local stack. The given worlds $W_3$ and $W_5$ are public
-future worlds in which the adversary chooses to return to us. None of these
-worlds can have replaced the temporary regions of $W_1$ with permanent regions.
-Therefore, by revoking the temporary regions of $W_5$ and reinstating the
-temporary regions of $W_1$, we can construct $W_6$ as a private future world of
-$W_5$ but also a public future world of $W_1$. Because of the latter, it is safe
-to use the local return pointer to return to the adversary, since we have
-restored validity of any local state they might have stashed away.
+first with a capability for the closure creator \texttt{\footnotesize{g1}}. Consequently, we need
+to check that all arguments are safe to use and that we clean up before
+returning in the end. The inversion of control poses an interesting challenge
+when it comes to reasoning about the adversary's local state during the
+execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
+the adversary should not be able to rely on its local state from before the call
+of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
+world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
+adversary is again allowed to rely on its old local state so we
+need to guarantee that the local state is unchanged. This is important because
+the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
+and the adversary is allowed to allocate the activation record on the stack
+(just like we do) so they can store and recover their old stack pointer after
+\texttt{\footnotesize{f4}} returns.
+
+In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in the
+proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
+all the situations where control is passed to the adversary. The potentially
+local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
+$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
+constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
+by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
+we pass away and static temporary regions (a region that accepts exactly one
+memory) for our local stack as well as the adversary's local stack. The given
+worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
+to return to us. None of these worlds can have masked the temporary regions of
+$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
+regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
+$W_6$ a public future world of $W_1$. It is therefore safe to use the return
+pointer to return to the adversary, since we have restored validity of any local
+state they might have stashed away.
 
@@
 unstorable status of evaluation contexts in lambda calculi without control
-effects. To reason about code that relies on this (particularly, the original
-awkward example), \citet{Dreyer:jfp12} (DNB) formally capture the special status
-of evaluation contexts using Kripke worlds with public and private future world
-relations. Essentially, they allow relatedness of evaluation contexts to be
-monotone with respect to a weaker future world relation (public) than
-relatedness of values. This formalizes the idea that it is safe to make
-temporary internal state modifications (private world transitions, that
-invalidate the continuation, but not other values) while a term is performing
-internal steps, as long as it returns to a stable state (i.e. transitions to a
-public future world of the original) before returning. We generalize this idea
-to reason about local capabilities: validity of local capabilities is allowed to
-be monotone with respect to a weaker future-world relation than other values,
-which we can exploit to distinguish between state changes that are always valid
-(public future worlds) and changes that are only valid if we clear local
-capabilities (private future worlds). Our future world relations are similar to
-\citet{Dreyer:jfp12}'s (for example, our proof of the awkward example uses the
-exact same state transition system), but they turn up in an entirely different
-place in the logical relation: rather than using public future worlds for the
-special syntactic category of evaluation contexts, they are used in the value
-relation depending on the locality of the capability at hand. Additionally, our
-worlds are more complex because, to allow local memory capabilities and
-write-local capabilities, they can contain (revokable) temporary regions that
-are only monotonous w.r.t. public future worlds, while DNB's worlds are entirely
+effects (of which well-bracketed control flow is a consequence). To reason about
+code that relies on this (particularly, the original awkward example),
+\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
+contexts using Kripke worlds with public and private future world relations.
+Essentially, they allow relatedness of evaluation contexts to be monotone with
+respect to a weaker future world relation (public) than relatedness of values,
+formalising the idea that it is safe to make temporary internal state
+modifications (private world transitions, which invalidate the continuation,
+but not other values) while an expression is performing internal steps, as long
+as the code returns to a stable state (i.e. transitions to a public future world
+of the original) before returning (i.e.
+invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation).
+Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
+We
+generalize this idea to reason about local capabilities: validity of local
+capabilities is allowed to be monotone with respect to a weaker future-world
+relation than other values, which we can exploit to distinguish between state
+changes that are always safe (public future worlds) and changes that are only
+valid if we clear all local capabilities (private future worlds). Our future
+world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
+the awkward example uses exactly the same state transition system),
+but they turn up in an entirely different place in the logical relation: rather
+than using public future worlds for the special syntactic category of evaluation
+contexts, they are used in the value relation depending on the
+locality of the capability at hand. Additionally, our worlds are a bit more
+complex because, to allow local memory capabilities and write-local
+capabilities, they can contain (revokable) temporary regions that are only
+monotonous w.r.t. public future worlds, while DNB's worlds are entirely
 permanent.

### 2017-02-16 csf-2017/paper.tex
style

@@
 respect to a weaker future world relation (public) than relatedness of values,
-formalising the idea that it is safe to make temporary internal state
+formalizing the idea that it is safe to make temporary internal state
 modifications (private world transitions, which invalidate the continuation,

### 2017-02-16 csf-2017/paper.tex
Almost at the page limit.

--- 2017-02-16 csf-2017/paper.tex
@@
 \figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics for a
-number of representative instructions. Essentially, a configuration $\Phi$ will
-either decode and execute the instruction at $\memreg(\pcreg)$ if it is
-executable and the current address is in the valid range or fail otherwise. The
-table in the figure shows for a given instruction $i$ the result of executing it
-in configuration $\Phi$. The $\fail$ and $\halt$ instructions obviously fail and
-halt respectively. The $\move{}{}$ instruction simply modifies the register file
-as requested and updates the $\pcreg$ to the next instruction using the
-meta-function $\stdUpdatePc{}$.
+few representative instructions. Essentially, a configuration $\Phi$ either
+decodes and executes the instruction at $\memreg(\pcreg)$ if it is executable
+and its address is in the valid range or otherwise fails. The table in the
+figure shows for instructions $i$ the result of executing them in configuration
+$\Phi$. $\fail$ and $\halt$ obviously fail and halt respectively. $\move{}{}$
+simply modifies the register file as requested and updates the $\pcreg$ to the
+next instruction using the meta-function $\stdUpdatePc{}$.
 
 The $\load{}{}$ instruction loads the contents of the requested memory location
-into the appropriate register, but only if the capability has appropriate
-authority (i.e.\ read permission and an appropriate range). The $\restricttwo{}{}$
-instruction will update a capability's permissions and global/local tag in
-the register file, but requires that the new permissions are weaker than the
-original and does not turn local capabilities into global ones. The $\geta{}{}$
-instruction queries the current address of a capability and stores it in another
-register.
+into a register, but only if the capability has appropriate authority (i.e.\
+read permission and an appropriate range). $\restricttwo{}{}$ updates a
+capability's permissions and global/local tag in the register file, but only if
+the new permissions are weaker than the original. It also never turns local
+capabilities into global ones. $\geta{}{}$ queries the current address of a
+capability and stores it in a register.
 
@@
 after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
-cannot be used to read, write or execute and their current address and range
-cannot be modified. They can only be used to jump to, but when that happens,
-their permission is updated to $\exec$. This special behavior makes them
-suitable for representing a kind of closures: an opaque package containing a
-piece of code together with local encapsulated state. Such a package can be
-built as an enter capability $c = ((\entry,\gl),\start,\addrend,\addr)$ where
-the range $[\start,\addr-1]$ contains local state (data or capabilities) and
-$[\addr,\addrend]$ contains instructions. The package is opaque to an adversary
-holding $c$ but when $c$ is jumped to, the instructions can start executing and
-have access to the local data through the updated version of $c$ that is then in
-$\pcreg$\footnote{Note that the capability in the $\pcreg$-register can be used
-  to load from.}.
+cannot be used to read, write or execute and their address and range cannot be
+modified. They can only be used to jump to, but when that happens, their
+permission changes to $\exec$. They can be used to represent a kind of closures:
+an opaque package containing a piece of code together with local encapsulated
+state. Such a package can be built as an enter capability $c =
+((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
+contains local state (data or capabilities) and $[\addr,\addrend]$ contains
+instructions. The package is opaque to an adversary holding $c$ but when $c$ is
+jumped to, the instructions can start executing and have access to the local
+data through the updated version of $c$ that is then in $\pcreg$\footnote{Note
+  that the capability in the $\pcreg$-register can be used to load from.}.
 % \lau{As we

### 2017-02-16 csf-2017/paper.tex
Even closer to the page limit.

@@
 
-The \emph{register-file relation} $\stdrr$ specifies that register-files are
-safe when they contain safe words (i.e.\ words in $\stdvr$) in all registers but
-$\pcreg$. The \emph{expression relation} $\stder$ defines that a word is safe to
-use as a program counter if it can be plugged into a safe register file (i.e.\ a
-register file in $\stdrr$) and paired with a memory satisfying the world to
-produce a safe configuration.
+The \emph{register-file relation} $\stdrr$ defines safe register-files as those
+that contain safe words (i.e.\ words in $\stdvr$) in all registers but $\pcreg$.
+The \emph{expression relation} $\stder$ defines that a word is safe to use as a
+program counter if it can be plugged into a safe register file (i.e.\ a register
+file in $\stdrr$) and paired with a memory satisfying the world to produce a
+safe configuration.
 

### 2017-02-16 csf-2017/paper.tex
adverb used as adjective/noun
Shorten summary table to reach page limit.

@@
 \textbf{Calling convention} With the security measures introduced and motivated,
-here is a summary of our proposed calling convention:
+let us summarize our proposed calling convention:
 \begin{description}[font=\normalfont\itshape]
-\item[At program start-up] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system never hands out global write-local capabilities.
-\item[Before returning to the adversary] clear the non-return-value registers and the entire part of the stack that we had access to.
-\item[Before invoking the adversary] push an activation record to the stack, create a return pointer as a local $\entry$-capability to the instructions in the activation record, restrict the stack capability to the unused part in $r_\stk$, clear that part of the stack, and clear the non-argument registers.
-\item[Before invoking an adversary callback] make sure the callback capability is global.
-\item[When called by an adversary] make sure the received stack pointer has permission $\rwlx$.
+\item[At program start-up] A local $\rwlx$ stack pointer in register $r_\stk$. No global write-local capabilities.
+\item[Before returning to the adversary] Clear non-return-value registers. Clear the part of the stack we had access to.
+\item[Before invoking the adversary] Push activation record to the stack. Create
+  return pointer as local $\entry$-capability to the instructions in the record.
+  Restrict the stack capability to the unused part and clear it. Clear
+  non-argument registers.
+\item[Before invoking an adversary callback] Make sure callback is global.
+\item[When called by an adversary] Make sure received stack pointer has permission $\rwlx$.
 \end{description}

### 2017-02-16 csf-2017/paper.tex

@@
 \caption{Illustration of the worlds in the proof of
-  Lemma~\ref{lem:correctness-g1}. In the proof, the above row of worlds are
-  constructed by us, while the below worlds are given. $W_6$ is constructed such
+  Lemma~\ref{lem:correctness-g1}. In the proof, the top row of worlds are
+  constructed by us, while the bottom row of worlds are given. $W_6$ is constructed such
   that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
@@
 unstorable status of evaluation contexts in lambda calculi without control
-effects. To reason about code that relies on this (particularly, the original
-awkward example), \citet{Dreyer:jfp12} (DNB) formally capture the special status
-of evaluation contexts using Kripke worlds with public and private future world
-relations. Essentially, they allow relatedness of evaluation contexts to be
-monotone with respect to a weaker future world relation (public) than
-relatedness of values. This formalizes the idea that it is safe to make
-temporary internal state modifications (private world transitions, that
-invalidate the continuation, but not other values) while a term is performing
-internal steps, as long as it returns to a stable state (i.e. transitions to a
-public future world of the original) before returning. We generalize this idea
-to reason about local capabilities: validity of local capabilities is allowed to
-be monotone with respect to a weaker future-world relation than other values,
-which we can exploit to distinguish between state changes that are always valid
-(public future worlds) and changes that are only valid if we clear local
-capabilities (private future worlds). Our future world relations are similar to
-\citet{Dreyer:jfp12}'s (for example, our proof of the awkward example uses the
-exact same state transition system), but they turn up in an entirely different
-place in the logical relation: rather than using public future worlds for the
-special syntactic category of evaluation contexts, they are used in the value
-relation depending on the locality of the capability at hand. Additionally, our
-worlds are more complex because, to allow local memory capabilities and
-write-local capabilities, they can contain (revokable) temporary regions that
-are only monotonous w.r.t. public future worlds, while DNB's worlds are entirely
+effects (of which well-bracketed control flow is a consequence). To reason about
+code that relies on this (particularly, the original awkward example),
+\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
+contexts using Kripke worlds with public and private future world relations.
+Essentially, they allow relatedness of evaluation contexts to be monotone with
+respect to a weaker future world relation (public) than relatedness of values,
+formalizing the idea that it is safe to make temporary internal state
+modifications (private world transitions, which invalidate the continuation,
+but not other values) while an expression is performing internal steps, as long
+as the code returns to a stable state (i.e. transitions to a public future world
+of the original) before returning (i.e.
+invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation).
+Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
+We
+generalize this idea to reason about local capabilities: validity of local
+capabilities is allowed to be monotone with respect to a weaker future-world
+relation than other values, which we can exploit to distinguish between state
+changes that are always safe (public future worlds) and changes that are only
+valid if we clear all local capabilities (private future worlds). Our future
+world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
+the awkward example uses exactly the same state transition system),
+but they turn up in an entirely different place in the logical relation: rather
+than using public future worlds for the special syntactic category of evaluation
+contexts, they are used in the value relation depending on the
+locality of the capability at hand. Additionally, our worlds are a bit more
+complex because, to allow local memory capabilities and write-local
+capabilities, they can contain (revokable) temporary regions that are only
+monotonous w.r.t. public future worlds, while DNB's worlds are entirely
 permanent.

### 2017-02-16 csf-2017/paper.tex
Refix something I thought I'd fixed before.

@@
 as the code returns to a stable state (i.e. transitions to a public future world
-of the original) before returning (i.e.
-invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation).
-Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
-We
+of the original) before returning. We
 generalize this idea to reason about local capabilities: validity of local

### 2017-02-16 csf-2017/paper.tex
minor errors

--- 2017-02-16 csf-2017/paper.tex
@@
 explicit. As future work, we consider looking at a more explicit and useful way
-to formalize control flow correctness. The idea would be to define a variant
+to formalize control-flow correctness. The idea would be to define a variant
 of our capability machine with call and return instructions and
@@
 store the old stack pointer, but the adversary's return pointers are also
-allowed to be local too. Adversary callbacks on the other hand are required to
+allowed to be local. Adversary callbacks on the other hand are required to
 be global but the callbacks we construct ourselves are allocated on the heap and

### 2017-02-16 csf-2017/paper.tex
added a small text about global rwlx caps. (and broke the page limit again).
shortening in examples section

@@
 technical requirement, defined in the technical appendix.
+An interesting property of the read and write condition is that they
+prohibit global write-local capabilities which, as discussed in
+Section~\ref{sec:stack-and-return-pointer}, is necessary for the safe
+use of local capabilities.
 
@@
 
-In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
-compared to the previous lemmas. In this lemma, the adversary assumes control
-first with a capability for the closure creator \texttt{\footnotesize{g1}}. Consequently, we need
-to check that all arguments are safe to use and that we clean up before
-returning in the end. The inversion of control poses an interesting challenge
-when it comes to reasoning about the adversary's local state during the
-execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
-the adversary should not be able to rely on its local state from before the call
-of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
-world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
-adversary is again allowed to rely on its old local state so we
-need to guarantee that the local state is unchanged. This is important because
-the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
-and the adversary is allowed to allocate the activation record on the stack
-(just like we do) so they can store and recover their old stack pointer after
+In Lemma~\ref{lem:correctness-g1} the order of control has been
+inverted compared to the previous lemmas. In this lemma, the adversary
+assumes control first with a capability for the closure creator
+\texttt{\footnotesize{g1}}. Consequently, we need to check that all
+arguments are safe to use and that we clean up before returning in the
+end. The inversion of control poses an interesting challenge when it
+comes to reasoning about the adversary's local state during the
+execution of \texttt{\footnotesize{f4}} and the callbacks where the
+adversary should not rely on the local state from before the call of
+\texttt{\footnotesize{f4}}. This is easily done by revoking all the
+temporary regions of the world given at the start of
+\texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}}
+returns, the adversary is again allowed to rely on its old local state
+so we need to guarantee that the local state is unchanged. This is
+important because the return pointer that \texttt{\footnotesize{f4}}
+receives may be local, and the adversary is allowed
+to allocate the activation record on the stack (just like we do) so
+they can store and recover their old stack pointer after
 \texttt{\footnotesize{f4}} returns.

### 2017-02-16 csf-2017/paper.tex
shortened public private future world discussion

@@
 \emph{Public/private future worlds} A novel aspect of our logical relation is
-how we model the special nature of local capabilities using public/private
-future worlds. The main insight is that the temporary, revokable nature of these
-capabilities can be seen as a generalization of the syntactically-enforced
+how we model the temporary, revokable nature of local capabilities using public/private
+future worlds. The main insight is that this special nature generalizes that of the syntactically-enforced
 unstorable status of evaluation contexts in lambda calculi without control

### 2017-02-16 csf-2017/paper.tex
Lars comments

@@
 back. In this case, the adversary could use the unused part of the stack to
-store local pointers and load them from there after he gets control back. To
+store local pointers and load them from there after they get control back. To
 prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
@@
 the allocated block and pointing to the contained prelude instructions. However,
-the question that remains in this setup is: where do we get a stack pointer when
+the question that remains in this setup is: where do we get a stack pointer from when
 the callback is invoked?

### 2017-02-16 csf-2017/paper.tex
more update based on Lars comments

@@
 is a faithful translation of a tricky example  known from the
-literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
+literature (known as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
 parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
@@
 
-%\bibliographystyle{IEEEtran}
 \bibliographystyle{plainnat}
+%\bibliographystyle{plain}
 \bibliography{IEEEabrv,references}

### 2017-02-16 csf-2017/paper.tex
Fix citations to use IEEE style and to use natbib properly everywhere.

@@
 \usepackage{stmaryrd}
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{url}
@@
 \usepackage{multicol}
-\usepackage{natbib}
 \lstset{
@@
 machine code rewriting
-(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
-verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a
+(e.g.,~\citet{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
+verification (e.g.,~\citet{morrisett_system_1999}), virtual machines with a
 native stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
@@
 the hardware level. They are based on old ideas
-(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
+(e.g.~\citet{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
 but have recently received renewed interest; in particular, the CHERI
@@
   guarantees offered by a capability machine (a form of capability
-  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
+  safety~\citet{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
   guarantees offered for local capabilities. It is very general and not tied to
@@
 For reasons of space, some details and all proofs have been omitted;
-refer to the technical appendix~\cite{technical_appendix} for those.
+refer to the technical appendix~\citep{technical_appendix} for those.
 
@@
 like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
-both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
-CHERI~\cite{Watson2015Cheri}. To avoid uninteresting details, we assume an
+both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and
+CHERI~\citep{Watson2015Cheri}. To avoid uninteresting details, we assume an
 infinite address space and unbounded integers.
@@
 but it is complicated by the presence of \emph{enter capabilities}, modeled
-after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
+after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
 cannot be used to read, write or execute and their address and range cannot be
@@
 instruction is complicated by the presence of \emph{local capabilities}, modeled
-after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
+after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
 capabilities are special in that they can only be kept in registers, i.e.\ they
@@
 certain specification. Malloc and linking tables are described further in the
-next section, but we refer to the technical appendix~\cite{technical_appendix}
+next section, but we refer to the technical appendix~\citep{technical_appendix}
 for full details.
@@
 
-The technical appendix~\cite{technical_appendix} contains detailed descriptions
+The technical appendix~\citep{technical_appendix} contains detailed descriptions
 of all the macros.
@@
 similarly to what has been used in logical relations for high-level
-languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
+languages~\citep{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
 imposed by permanent regions stay in place indefinitely. Any capability, local
@@
 paragraphs. We omit some definitions and explain them only verbally, but precise
-definitions can be found in the technical appendix~\cite{technical_appendix}.
+definitions can be found in the technical appendix~\citep{technical_appendix}.
 
@@
 failing when an unauthorized access is attempted, like loading from a capability
-without read permission. This is similar to \cite{Devriese:2016ObjCap}'s logical
+without read permission. This is similar to \citet{Devriese:2016ObjCap}'s logical
 relation for an untyped language, but unlike typical logical relations
@@
 in the lemmas we state here. The interested reader can find all the technical
-details in the technical appendix~\cite{technical_appendix}.
+details in the technical appendix~\citep{technical_appendix}.
 
@@
 Formulations in terms of a control-flow graph, like the one by
-\cite{abadi_control-flow_2005}, do not take into account temporal properties,
+\citet{abadi_control-flow_2005}, do not take into account temporal properties,
 like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
@@
 with recursive
-worlds~\cite{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
+worlds~\citep{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
 closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{IEEEtranN}
+% \bibliographystyle{plainnat}
 %\bibliographystyle{plain}

### 2017-02-16 csf-2017/paper.tex
Add a few instances of "The macro" to make the introduction of the macro-instruction more readable.

@@
 macros that uses the stack assumes a stack pointer in register
-$r_\stk$. \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
+$r_\stk$. The macro \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
 capability related to $\var{name}$ from the linking table and stores
-it in register $r$.  \texttt{\footnotesize{push $r$}} and
+it in register $r$.  The macros \texttt{\footnotesize{push $r$}} and
 \texttt{\footnotesize{pop $r$}} work as one would
-expect. \texttt{\footnotesize{prepstk $r$}} is used when a callback is
+expect. The macro \texttt{\footnotesize{prepstk $r$}} is used when a callback is
 invoked by the adversary and prepares the received stack pointer by
-checking that it has permission $\rwlx$.  \texttt{\footnotesize{scall
+checking that it has permission $\rwlx$.  The macro \texttt{\footnotesize{scall
     $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}}
@@
 the stack, clears the necessary registers and jumps to $r$.  Upon
-return, the private state is restored. \texttt{\footnotesize{mclear
+return, the private state is restored. The macro \texttt{\footnotesize{mclear
     $r$}} clears all the memory the capability in register $r$ has
-authority over.  \texttt{\footnotesize{rclear $\var{regSet}$}} clears
-all the registers in $\var{regSet}$. \texttt{\footnotesize{reqglob
+authority over.  The macro \texttt{\footnotesize{rclear $\var{regSet}$}} clears
+all the registers in $\var{regSet}$. The macro \texttt{\footnotesize{reqglob
     $r$}} checks whether the word in register $r$ is a global
-capability. \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}}
+capability. The macro \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}}
 allocates a closure where $r$ points to the closure's code and a new

### 2017-02-16 csf-2017/paper.tex
Readd the reference to the categorical logic tutorial notes.

@@
 world-dependent. Recursive worlds are a common requirement in Kripke models and
-the following lemma uses the method of \citet{Birkedal:2011:SKM:1926385.1926401}
-for constructing them. The formulation of the lemma is technical, so we
-recommend that non-expert readers ignore the technicalities and accept that
-there exists a set of worlds $\Wor$ and two relations $\futurestr$ and
-$\futurewk$ satisfying the (recursive) equations in the theorem (where the
-$\blater$ operator can be safely ignored).
+the following lemma uses the method of
+\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
+constructing them. The formulation of the lemma is technical, so we recommend
+that non-expert readers ignore the technicalities and accept that there exists a
+set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying
+the (recursive) equations in the theorem (where the $\blater$ operator can be
+safely ignored).
 % Short technical explanation

### 2017-02-17 csf-2017/paper.tex
DNB fix

@@
 valid if we clear all local capabilities (private future worlds). Our future
-world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
+world relations are similar to DNB's (for example, our proof of
 the awkward example uses exactly the same state transition system),

### 2017-02-17 local-capability-machine.tex
Took a pass over the proof of lemma 3.

@@
     So it suffices to show \ref{pf:f2-sc:sfts-obs}. To this end use Lemma~\ref{lem:anti-red-obs}. Let $\ms_f$ be given, then
-  \begin{multline*}
-    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},\ms \uplus \ms_\stk \uplus \ms_f) \step[k] \\
+    \[
+    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},\ms \uplus \ms_\stk \uplus \ms_f) \step[k] 
     (\reg',\ms \uplus \ms_\stk' \uplus \ms_f)
-  \end{multline*}
+  \]
   where
@@
   \begin{enumproof}
-  \item $\memSat{n-k}{\ms}{W}$\\
+  \item $\memSat[n-k]{\ms}{W}$\\
     Use Lemma~\ref{lem:disj-mem-sat} with
@@
 \end{lemma}
+In an attempt to aid the reader, we first provide to high-level descriptions of possible proof of Lemma~\ref{lem:correctness-f3} followed by a more detailed proof.
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3} (high-level description)]
@@
           & [1 \mapsto \iota^\sta (\perma,\ms_\flag \uplus \ms_{f2})]\\
-          & [2 \mapsto \iota^\sta (\perma,\ms_\link)]\\
+          & [2 \mapsto \iota^{\sta,u} (\perma,\ms_\link)]\\
           & [3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]\\
@@
 then using $\ms_{\var{frame}}$ as the frame and $m'$ as the resulting memory, we get that $m' = \ms'' \uplus \ms_r \uplus \ms_{\var{frame}}$ for some $\ms'$ and $\ms_r$ s.t.
-$\heapSat[1]{\ms''}{W}$.
+$\memSat[1]{\ms''}{W}$.
 Region $1$ guarantees that the assertion flag is unchanged, so we have
@@
 \]
-where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction, namely \texttt{pop r1}.
-Further we have \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk |_{\start_\stk+1,\addrend_\stk}} and $i$ is a suitable number of steps.
+where $c_{\var{next}}$ is $c_{f3}$ adjusted to point to the next instruction, namely \texttt{pop r1}. Further we have
+\begin{itemize}
+\item \pointstostack{\reg'}{[\start_\stk \mapsto 1]}{\ms_\stk|_{\start_\stk+1,\addrend_\stk}}
+\end{itemize}
+and $i$ is a suitable number of steps.
 
@@
 \]
-We use Lemma~\ref{lem:scall-works} which requires us to show
+We use Lemma~\ref{lem:scall-works} (we do not use the local frame in the lemma) which requires us to show
 \begin{enumproof}
@@
       \item $\ms_\flag \uplus \ms_{f2}$: governed by region 1, only this memory segment is accepted.
-      \item $\ms_\link$: governed by region 2, only this memory segment is accepted.
-      \item $\ms_\adv$:\\
-        show
-        \[
-          \npair[n]{\ms_\adv} \in H^\pwl_{\start_\adv,\addrend_\adv}\;1\;\xi^{-1}(W)
-        \]
-        this is trivial except for showing
-        \[
-          \npair[n-1]{c_\link} \in \stdvr(W)
-        \]
-        amounts to showing
-        \[
-          \npair[n-1]{\link,\link+1} \in \readCond{}(\glob)(W)
-        \]
-        which follows from
-        \begin{enumproof}
-          \item $\npair[n-1]{c_\adv} \in \stdvr(W)$: \label{item:f3:adv1}\\
-            We will show
-            \begin{equation}
-              \label{eq:f3:adv-in-stder}
-              \forall W' \futurestr W \ldotp \npair{c_\adv} \in \stdvr(W')
-            \end{equation}
-            which will give us \ref{item:f3:adv1} using downwards closure.
-
-            Let $W' \futurestr W$ be given and show
-            \[
-              \npair{(\start_\adv,\addrend_\adv,\start_\adv+\olf)} \in \entryCond{}(\glob)(W')
-            \]
-            to this end let $W'' \futurestr W'$ and $n' < n$ be given and show
-            \[
-              \npair[n']{\updatePcPerm{c_\adv}} \in \stder(W'')
-            \]
-            This follows from the FTLR if we can show
-            \[
-              \npair[n']{\start_\adv,\addrend_\adv} \in \readCond{}(\glob)(W'')
-            \]
-            $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}$ governs the adversary, so the result follows from Lemma~\ref{lem:nwlp-subset-pwl}.
-          \item $\npair[n-1]{c_\malloc} \in \stdvr(W)$\\
-            Which follows from Lemma~\ref{lem:malloc-in-vr}.
-            \lau{I don't think this is actually the case at the moment. I guess we either need to show this based on the specification or just require it in the specification.}
-        \end{enumproof}
+      \item $\ms_\link$: governed by region 2, only this memory segment is accepted. We also need to show that the contents is safe, i.e. shoe
+        \begin{enumproof}
+          \item $\npair[n-i]{c_\malloc} \in \stdvr(W)$: Follows from Lemma~\ref{lem:malloc-in-vr}.
+          \item $\npair[n-i]{c_\adv} \in \stdvr(W)$: Follows from Theorem~\ref{thm:ftlr} and Lemma~\ref{lem:nwlp-subset-pwl}.
+        \end{enumproof}
+      \item $\ms_\adv$:
+        Follows from Lemma~\ref{lem:stau-subset-pwl}.
     \end{enumproof}
@@
     \item $\memSat[n-i-1]{\ms''}{W'}$
-    \item $\reg'' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}'}$ unused}$
-    \item $\forall a \in \dom(\ms_{\var{unused}}') \ldotp \ms''(a) = 0$
-    \item $\reg''= \reg_0[\pcreg\mapsto\updatePcPerm{\reg'(r)},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto \_, r \mapsto \reg'(r)]$ 
+    \item $\reg'' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
+    \item $\reg''= \reg_0[\pcreg\mapsto\updatePcPerm{\reg'(r)},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk', r \mapsto \reg'(r)]$ 
     \item $\npair[n-i-1]{c_{\mathit{ret}}} \in \stdvr(W')$
+    \item $\npair[n-i-1]{c_\stk'} \in \stdvr(W')$
     \end{itemize}
@@
           \item $\npair[n-i-1]{c_{\mathit{ret}}} \in \stdvr(W')$ : by above assumptions.
-          \item $\npair[n-i-1]{reg''(r_\stk)} \in \stdvr(W')$ : by Lemma~\ref{lem:stack-cap-vr}.
+          \item $\npair[n-i-1]{c_\stk'} \in \stdvr(W')$ : by above assumptions.
           \item $\npair[n-i-1]{c_\adv} \in \stdvr(W')$ : follows from \ref{eq:f3:adv-in-stder}.
@@
     \item $n' \leq n-i-2$
-    \item $W'' \futurewk \revokeTemp{W}[\iota^{\sta}(\temp,[\start_\stk \mapsto 1])[\iota^\sta(\temp,\ms_{\mathit{unused}}'')]$
+    \item $W'' \futurewk \revokeTemp{W}$
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
@@
       \begin{enumproof}
-      \item $\npair[n''-k]{(\reg^{(3)},\ms'' \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}'')} \in \observations(\revokeTemp{W''})$\\
-        Show this using Lemma~\ref{lem:scall-works}. Show:
-        \begin{enumproof}
-          \item $\memSat[n''-k]{\ms''}{\revokeTemp{W''}}$ is satisfied by one of the first Hyp-cont assumptions.
+      \item $\npair[n'-k]{(\reg^{(3)},\ms'' \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}'')} \in \observations(\revokeTemp{W''})$\\
+        Show this using Lemma~\ref{lem:scall-works} a. Show:
+        \begin{enumproof}
+          \item $\memSat[n'-k]{\ms''}{\revokeTemp{W''}}$ is satisfied by one of the first Hyp-cont assumptions and Lemma~\ref{lem:heap-sat-dc}.
           \item Hyp-Callee\\
@@
             \item $W^{(3)} = \revokeTemp{W''}[\iota^{\sta}(\temp,[\start_\stk \mapsto 2]\uplus\ms_{\mathit{act}}'),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}^{(3)}))]$
-            \item $\memSat[n''-k-1]{\ms^{(3)}}{W^{(3)}}$
+            \item $\memSat[n'-k-1]{\ms^{(3)}}{W^{(3)}}$
             \item $\reg^{(4)} \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}^{(3)}$ unused}$
-            \item $\forall a \in \dom(\ms_{\var{unused}}^{(3)}) \ldotp \ms^{(3)}(a) = 0$
-            \item $\reg^{(4)}= \reg_0[\pcreg\mapsto\updatePcPerm{c_\adv},r_0 \mapsto c_{\mathit{ret}}', r_\stk \mapsto \_, r \mapsto c_\adv]$ 
-            \item $(n''-k-1,c_{\mathit{ret}}') \in \stdvr(W')$
+            \item $\reg^{(4)}= \reg_0[\pcreg\mapsto\updatePcPerm{c_\adv},r_0 \mapsto c_{\mathit{ret}}', r_\stk \mapsto c_\stk'', r \mapsto c_\adv]$ 
+            \item $(n'-k-1,c_{\mathit{ret}}') \in \stdvr(W^{(3)})$
+            \item $(n'-k-1,c_\stk'') \in \stdvr(W^{(3)})$
             \end{itemize}
@@
             \begin{enumproof}
-              \item $\memSat[n''-k-1]{\ms^{(3)}}{W^{(3)}}$: By assumption.
-              \item $\npair[n''-k-1]{\reg^{(4)}}$: Show:
-                \begin{enumproof}
-                  \item $\npair[n''-k-1]{c_\adv} \in \stdvr(W^{(3)})$ by \ref{eq:f3:adv-in-stder}.
-                  \item $\npair[n''-k-1]{c_{\mathit{ret}}'}$ by assumption.
-                  \item $\npair[n''-k-1]{\reg^{(4)}(r_\stk)}$ by Lemma~\ref{lem:stack-cap-vr}.
+              \item $\memSat[n'-k-1]{\ms^{(3)}}{W^{(3)}}$: By assumption.
+              \item $\npair[n'-k-1]{\reg^{(4)}} \in \stdrr(W^{(3)})$: Show:
+                \begin{enumproof}
+                  \item $\npair[n'-k-1]{c_\adv} \in \stdvr(W^{(3)})$ by Assumption \ref{eq:f3:adv-in-stder}.
+                  \item $\npair[n'-k-1]{c_{\mathit{ret}}'}$ by assumption.
+                  \item $\npair[n'-k-1]{c_\stk''}$ by assumption
                 \end{enumproof}
@@
             \[
-              \npair[n''-k-1]{(\reg^{(4)},\ms^{(3)})} \in \observations(W^{(3)})
+              \npair[n'-k-1]{(\reg^{(4)},\ms^{(3)})} \in \observations(W^{(3)})
             \]
@@
             \begin{itemize}
-            \item $n^{(3)} \leq n''-k-2$
+            \item $n'' \leq n'-k-2$
             \item $W^{(3)} \futurewk \revokeTemp{W''}[\iota^{\sta}(\temp,\ms_\stk)][\iota^\sta(\temp,\ms_{\mathit{unused}}^{(3)})]$
-            \item $\memSat[n^{(3)}]{\ms^{(3)}}{\revokeTemp{W^{(3)}}}$ 
+            \item $\memSat[n'']{\ms^{(3)}}{\revokeTemp{W^{(3)}}}$ 
             \item for all $r$, we have that:
@@
             \[
-              (n^{(3)}, (\reg^{(3)},\ms^{(3)} \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}^{(3)})) \in \observations(\revokeTemp{W^{(3)}})
-            \]
-            To this end let $\ms_f''$, $m''$, and $j \leq n^{(3)}$ be given and assume
+              (n'', (\reg^{(3)},\ms^{(3)} \uplus [\start_\stk \mapsto 2] \uplus \ms_{\var{unused}}^{(3)})) \in \observations(\revokeTemp{W^{(3)}})
+            \]
+            To this end let $\ms_f''$, $m''$, and $j \leq n''$ be given and assume
             \[
@@
             \]
-            By assumption we had $\memSat[n^{(3)}]{\ms^{(3)}}{\revokeTemp{W^{(3)}}}$ and the frame is unchanged, so we can split the memory as desired.
-        \end{enumproof}
-      \end{enumproof}
-    \end{enumproof}
-  \end{enumproof}
-\end{proof}
-
+            By assumption we had $\memSat[n'']{\ms^{(3)}}{\revokeTemp{W^{(3)}}}$ and the frame is unchanged, so we can split the memory as needed.
+        \end{enumproof}
+      \end{enumproof}
+    \end{enumproof}
+  \end{enumproof}
+\end{proof}
+
+\begin{comment}
 \begin{proof}[Proof of Lemma~\ref{lem:correctness-f3}]
@@
 \end{proof}
+\end{comment}
 

### 2017-02-17 local-capability-machine.tex
pass over the entire example section to remove things we do not use anymore

@@
               \subsection{Labels}
-              \texttt{l:} is a meta level label that can be used to refer to a specific address.
+              \texttt{l:} is a meta level label that can be used to refer to a specific address. When placed on the line of a macro, it refers to the first instruction of this macro.
 
@@
               \label{sec:examples}
-              \subsection{Simple Examples Using Local Capabilities}
-              \label{subsec:example-loc-cap}
-              \lau{This text regards the ML-like programs and can be removed when they are.}
-              In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1.
+%               \lau{This text regards the ML-like programs and can be removed when they are.}
+%               In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1.
 The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}).
 Notice that an assertion will never terminate in a $\failed$ state.
-              \subsubsection{Simple use of local capabilities}
-              The following programs only use local capabilities when they create protected return pointers.
+%               \subsubsection{Simple use of local capabilities}
+%               The following programs only use local capabilities when they create protected return pointers.
 The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
 
-              ML-like program we had before:
-\begin{verbatim}
-let f = fun adv =>
-          let l = 1 in
-          adv();
-          assert (l == 1)
-\end{verbatim}
-              \lau{The above program is obsolete (and the other ML-like programs), but I will not remove it until we have gotten used to the assembly programs.}
-
+%               ML-like program we had before:
+% \begin{verbatim}
+% let f = fun adv =>
+%           let l = 1 in
+%           adv();
+%           assert (l == 1)
+% \end{verbatim}
+%               \lau{The above program is obsolete (and the other ML-like programs), but I will not remove it until we have gotten used to the assembly programs.}
+
+\subsection{Encapsulation of Local State}
 Assembly program not using stack. Assume that $\mathtt{r_l} \not\in \{\pcreg,r_0 \}$ is a register.
@@
 \end{verbatim}
-
+For \texttt{f1} to work, its local state needs to be encapsulated. 
 \begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
@@
 
+\subsection{Encapsulation of Local State Using Local Capabilities and \texttt{scall}}
+\label{subsec:example-loc-cap}
+
 Assembly program using the stack. This program assumes a $r_\stk \not\in \{\pcreg,r_0\}$ register that contains a stack capability (a local $\rwlx$-capability):
@@
 
-\subsection{Simple well-bracketedness/local capability example}
+\subsection{Well-Bracketedness Using Local Capabilities and \texttt{scall}}
 \begin{comment}
@@
 
-\subsection{Another simple well-bracketedness/local capability example}
-Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability (but \texttt{let x = ...} produces a global capability). 
-\begin{verbatim}
-let f = fun adv =>
-          let x = 1 in
-          let l = 0 in
-          adv(l);
-          assert(!x == 1);
-          x := 2;
-          adv(0);
-          return
-\end{verbatim}
-In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}.
+% \subsection{Another simple well-bracketedness/local capability example}
+% Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability (but \texttt{let x = ...} produces a global capability). 
+% \begin{verbatim}
+% let f = fun adv =>
+%           let x = 1 in
+%           let l = 0 in
+%           adv(l);
+%           assert(!x == 1);
+%           x := 2;
+%           adv(0);
+%           return
+% \end{verbatim}
+% In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}.
 \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.
 
-\subsection{Inverted control and return from closure}
-If we can prove the below example using the scall lemma, then we should also be able to prove the awkward example.
+\subsection{Inverted Control and Return From Closure}
+The following example is constructed to investigate the difficulties of preserving an adversary's local frame.
+There is no assertion as this is (slightly) beside the point.
+The lemma we would prove about this should look like Lemma~\ref{lem:correctness-g1}, but it is not state and proven here.
 \begin{lstlisting}
@@
 \subsection{Variant of the ``awkward'' example}
-Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
+Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566} which roughly was:
 \begin{verbatim}
@@
 \end{verbatim}
+Our translation of the example:
 \begin{lstlisting}

### 2017-02-17 local-capability-machine.tex
pass over awkward example proof

@@
     \end{enumproof}
-  \item Hyp-Callee\\
+  \item Hyp-Callee \\
     Assume
@@
   \label{lem:iotax-is-a-region}
-  The region in Definition~\ref{def:iotax-region} is a region.
+  Definition~\ref{def:iotax-region} defines a region.
 \end{lemma}
@@
   \end{equation}
-  then, the termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 4 is permanent, so
+  then the termination assumption gives us that part of $m$ satisfies a private future world of $W$. Region 4 is permanent, so
   \[
@@
   \]
-  So it suffices to show \ref{eq:f4:start-conf-in-obs}. 
+  So it suffices to show Eq. \ref{eq:f4:start-conf-in-obs}. 
   To this end use the FTLR to show $\npair{c_\adv} \in \stder(W)$, so show
@@
   \end{enumproof}
-  in conclusion $\npair{c_\adv} \in \stder(W)$. We get \ref{eq:f4:start-conf-in-obs} if we show the following
+  in conclusion $\npair{c_\adv} \in \stder(W)$. We get Eq. \ref{eq:f4:start-conf-in-obs} if we show \ref{pf:g1-3} and \ref{pf:g1-4}:
   \begin{enumproof}[resume]
-    \item $\memSat{\ms \uplus \ms_\stk}{W}$\\
-      \begin{enumproof}
-        \item $\memSat{\ms_{g1} \uplus \ms_\flag}{[4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]}$ \\
+    \item $\memSat{\ms \uplus \ms_\stk}{W}$ \label{pf:g1-3}\\
+      \begin{enumproof}
+        \item $\memSat{\ms_{g1} \uplus \ms_\flag}{[4 \mapsto \iota^\sta (\perma,\ms_{g1} \uplus \ms_\flag)]}$ \label{pf:g1-3-1}\\
           Lemma~\ref{lem:mem-sat-static}.
-        \item $\memSat{\ms_\stk}{[2 \mapsto \iota^\pwl_{\start_\stk,\addrend_\stk}]}$\\
+        \item $\memSat{\ms_\stk}{[2 \mapsto \iota^\pwl_{\start_\stk,\addrend_\stk}]}$ \label{pf:g1-3-2}\\
           Lemma~\ref{lem:mem-sat-data-only-std-regions} and assumption that $\ms_\stk$ is all 0.
-        \item $\memSat{\ms_{\malloc} \uplus \ms_\link \uplus \ms_\adv}{[0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]}$\\
+        \item $\memSat{\ms_{\malloc} \uplus \ms_\link \uplus \ms_\adv}{[0 \mapsto \iota_{\malloc,0}][1 \mapsto \iota^{\sta,u} (\perma,\ms_\link)][3 \mapsto \iota^{\nwl,p}_{\start_\adv,\addrend_\adv}]}$ \label{pf:g1-3-3}\\
           For convenience define
@@
 
-          The memory link amounts to showing:
+          The linking table part of memory amounts to showing:
           \[
@@
           \]
-          which follows from Lemma~\ref{lem:malloc-in-vr}
+          which follows from Lemma~\ref{lem:malloc-in-vr}.
 
@@
       \end{enumproof}
-      Using Lemma~\ref{lem:disj-mem-sat} repeatedly with the above gives the desired memory satisfaction.
-    \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$\\
+      Using Lemma~\ref{lem:disj-mem-sat} repeatedly with \ref{pf:g1-3-1}, \ref{pf:g1-3-2}, and \ref{pf:g1-3-3} gives the desired memory satisfaction.
+    \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}} \in \stdrr(W)$ \label{pf:g1-4}\\
       This amounts to showing
@@
           \end{itemize}
-          and further there is a $\iota^\pwl$ region for $\ms_\stk$ in $W$, so the result follows from Lemma~\ref{stack-cap-vr}.
+          and further there is a $\iota^\pwl$ region for $\ms_\stk$ in $W$, so the result follows from Lemma~\ref{lem:stack-cap-vr}.
         \item $\npair{c_{g1}} \in \stdvr(W)$ \\
@@
           \[
-            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms^\env \uplus \ms^\cls)}\update{i_2}{\iota_x}
+            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms_\env \uplus \ms_\cls)}\update{i_2}{\iota_x}
           \]
@@
           \begin{enumproof}
-            \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls}{W_2}$
-              % TODO follows from \ms_1 heap sat and the new regions.
+            \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls}{W_2}$\\
+              We first show the following:
+              \begin{itemize}
+              \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{0}{\iota_\malloc}}$: we already know this.
+              \item $\memSat[n_2-i]{\ms_\env \uplus \ms_\cls}{[i_1 \mapsto \iota^\sta(\perma, \ms_\env \uplus \ms_\cls)]}$: By Lemma~\ref{lem:mem-sat-static}.
+              \item $\memSat[n_2-i]{\ms_x}{i_2 \mapsto \iota_x}$: $\ms(x) = 0$, so okay.
+              \end{itemize}
+
             \item $\npair[n_2-i]{\reg_2} \in \stdrr(W_2)$ \\
@@
                     \item $\memSat[n_4-j]{\ms_4}{\revokeTemp{W_4}}$: follows from $W_4 = \revokeTemp{W_4}$
-                    \item Hyp-Callee\\
+                    \item Hyp-Callee \label{pf:g1-first-hyp-callee}\\
                       We know $\npair[n_4]{\reg_3(r_1)} \in \stdvr(W_3)$.
@@
                           1) $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.
-
-                          2) The protected return pointer is in the value relation by Hyp-callee assumption, and the stack capability is handled by lemma~\ref{lem:stack-cap-vr}.
+                          2) The protected return pointer and the stack capability are in the value relation by Hyp-callee assumptions.
                       \end{enumproof}
@@
                       \item $n_5 \leq n_4-j-2$
-                      \item $W_6 \futurewk \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset \uplus \ms_\act),\iota^\pwl(\dom(\ms_\unused'')),\iota^\sta(\temp,\ms_f)]$ 
+                      \item $W_6 \futurewk \revokeTemp{W_4}$
                       \item $\memSat[n_5]{\ms_6}{\revokeTemp{W_6}}$
@@
           \item $\memSat[n_5-k]{\ms_6\update{x}{1}}{\revokeTemp{W_7}}$, follows from $W_7= \revokeTemp{W_7}$.
-          \item Hyp-Callee
+          \item Hyp-Callee: Goes like the first Hyp-Callee (\ref{pf:g1-first-hyp-callee}).
             % Use that it is the same $r_1$
@@
             \item $n_6 \leq n_5-k-2$
-            \item $W_8 \futurewk \revokeTemp{W_7}[\iota^{\sta}(\temp,\ms_\act)][\iota^\sta(\temp,\ms_{\mathit{unused}}^{(5)})][\iota^\sta(\temp,\ms_r)]$
+            \item $W_8 \futurewk \revokeTemp{W_7}$
             \item $\memSat[n_6]{\ms_7}{\revokeTemp{W_8}}$ 
@@
     &\dom(hs_\flag) = [\flag,\flag] \\
-    &\heapSat[\hs_{\adv} \uplus \hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}, 1 \mapsto\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}]}
+    &\heapSat[\hs_{\adv} \uplus \hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}, 1 \mapsto\iota^\nwl_{\start_\adv,\addrend_\adv}]}
   \end{align*}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:stder-ne}]
-
+  % TODO
 \end{proof}

### 2017-02-22 local-capability-machine.tex
added malloc and crtcls macro correctness lemmas.

@@
 \newcommand{\pointstostack}[3]{\ensuremath{#1} \text{ points to stack with } \ensuremath{#2} \text{ used and } \ensuremath{#3} \text{ unused}}
+\newcommand{\linksto}[4]{\ensuremath{#1} \text{ links } \ensuremath{#2} \text{ as } \ensuremath{#3} \text{ to } \ensuremath{#4}}
 \newcommand{\nonlocal}[1]{\ensuremath{#1} \text{ is non-local}}
@@
 
+              In order to compute offsets and the like, the macros need registers to keep temporary computations in.
+We assume such a small set of registers $\RegName_t \subseteq \RegName$ is available and that $\RegName_t$ does not contain registers explicitely named in a program nor $r_0$, $r_\stk,$ or $\pcreg$ (but clearing all registers still clears the temporary registers).
+
               \subsection{Linking and ABI}
@@
               \item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register $r$. 
+% Suggestion for malloc impl (r_t1 and r_t2 are registers in RegName_t), r_1 should not contain important stuff
+% move r_1 n
+% move r pc
+% getb r_t1 r
+% geta r_t2 r
+% minus r_t1 r_t2 r_t1
+% lea r r_t1
+% move r_0 r_t1
+% lea r_0 3 // 3 is offset to just after jmp
+% restrict r_0 encodePerm(e)
+% jmp r
+% move r r_1
+% move r_1 0
+% move r_0 r_t1
+% move r_t1 r_0
 
@@
 \subsection{Lemmas}
-
 \subsubsection{Anti-reduction for the observation relation}
@@
 
+\subsubsection{Malloc macro-instruction correctness}
+\begin{definition}
+  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c_\malloc}'' 
+  iff
+  \begin{itemize}
+  \item $\reg(pc) = \stdcap$
+  \item $\ms(\start) = ((\_,\_),\start_\link,\_,\_)$
+  \item $\ms(\start_\link+j) = c$
+  \end{itemize}
+\end{definition}
+
+\begin{lemma}[$\mathtt{malloc}$ works]
+  \label{lem:malloc-works}
+  If
+  \begin{itemize}
+  \item \lookingat{(\reg,\ms)}{\mathtt{malloc}\;r\;k}{c_{\var{next}}}
+  \item $k \geq 0$
+  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
+  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
+  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
+  \item $\memSat{\ms}{W}$
+  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
+  \item $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
+  \item[Hyp-Cont]
+    If
+    \begin{itemize}
+    \item $n' \leq n-1$
+    \item $\iota_\malloc \futurewk W(i)$
+    \item $\memSat[n']{\ms_{\var{footprint}}' \uplus \ms'}{W[i \mapsto \iota_\malloc]}$
+    \item $\memSat[n']{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
+      \begin{equation*}
+        \reg'(r') = 
+        \begin{cases}
+          c_{\var{next}} & r' = \pcreg \\
+          ((\rwx,\glob),\start,\addrend,\addr) & r' = r \\
+          \reg(r) & r' \not\in \RegName_t \union \{\pcreg, r, r_1\}
+        \end{cases}
+      \end{equation*}
+    \item $\addrend - \start = k - 1$
+    \item $\dom(\ms_{\var{alloc}}) = [\start,\addrend]$
+    \item $\forall \addr \in [\start,\addrend]\ldotp \ms_{\var{alloc}}(\addr) = 0$
+    \end{itemize}
+    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}}'  \uplus \ms_{\var{alloc}})} \in \observations(W[\iota_\malloc])$
+  \end{itemize}
+  Then
+  \[
+    \npair{(\reg,\ms)} \in \observations(W)
+  \]
+\end{lemma}
+
+\subsubsection{Create closure macro-instruction correctness}
+\begin{lemma}[$\mathtt{crtcls}$ works]
+  If
+  \begin{itemize}
+  \item \lookingat{(\reg,\ms)}{\mathtt{crtcls}\;\overline{(x,r)}\;r}{c_{\var{next}}}
+  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
+  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
+  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
+  \item $\memSat{\ms}{W}$
+  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
+  \item $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
+  \item[Hyp-Cont]
+    If
+    \begin{itemize}
+    \item $n' \leq n$
+    \item $\iota_\malloc \futurewk W(i)$
+    \item $\memSat[n']{\ms' \uplus \ms_{\var{footprint}}'}{W[i \mapsto \iota_\malloc]}$
+    \item $\memSat{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
+    \item
+      \begin{equation*}
+        \reg'(r') =
+        \begin{cases}
+          c_{\var{next}} & r' = \pcreg \\
+          c_\cls = ((\entry,\glob),\start,\addrend,\start+2) & r' = r_1 \\
+          \reg(r)      & r' \not\in \{\pcreg, r_1\} \union \RegName_t
+        \end{cases}
+      \end{equation*}
+    \item $\ms_\cls = \ms_\act \uplus \ms_\env$
+    \item $c_\env = ((\readwrite,\glob),\start_\env,\addrend_\env,\start_\env)$
+    \item $\dom(\ms_\env) = [\start_\env,\addrend_\env]$
+    \item $\ms_\env(\start_\env,\dots,\addrend_\env) = \reg(\overline{r})$
+    \item Hyp-act\\ If
+      \begin{itemize}
+      \item $\reg''(\pcreg) = \updatePcPerm{c_\cls}$
+      \end{itemize}
+      Then  $\exists k \ldotp \forall \ms_f \ldotp (\reg'',\ms'' \uplus \ms_\cls \uplus \ms_f) \step[k] (\reg''',\ms'' \uplus \ms_\cls \uplus \ms_f)$
+      where
+        \begin{equation*}
+          \reg'''(r') =
+          \begin{cases}
+            c_\env & r' = c_\env \\
+            \updatePcPerm{\reg(r)} & r' = \pcreg \\
+            \reg''(r') & r' \not\in \RegName_t
+          \end{cases}
+        \end{equation*}
+    \end{itemize}
+    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}} \uplus \ms_\cls)} \in \observations(W[i \mapsto \iota_\malloc])$
+  \end{itemize}
+  Then 
+  \[
+    \npair{(\reg,\ms)} \in \observations(W)
+  \]
+\end{lemma}
+
 \subsubsection{Helper lemmas about the stack}
@@
 \begin{proof}
-  By definition of $\stdrr(W)$ using Lemma~\ref{stdvr-dc}.
+  By definition of $\stdrr(W)$ using Lemma~\ref{lem:stdvr-dc}.
 \end{proof}

### 2017-02-23 local-capability-machine.tex
added comment with possible crtcls implementation
adjusted the g1 proof slightly

@@
                 \end{description}
+%Possible implementation of crtcls \overline{(x,r_v)} r_code where |\overline{(x,r_v)}| = n (i_1,...,i_6 under)
+% malloc r_t1 n
+% store r_t1 r_v1
+% lea r_t1 1
+% store r_t1 r_v2
+% lea r_t1 1
+% ...
+% lea r_t1 1
+% store r_t1 r_vn
+% lea r_t1 -n
+% restrict r_t1 encodePerm(rw)
+% malloc r_1 8 //length of activation record
+% store r_1 r_code
+% lea r_1 1
+% store r_1 r_t1
+% move r_t1 0
+% lea r_1 1
+% store r_1 encode(i_1)
+% lea r_1 1
+% store r_1 encode(i_2)
+% lea r_1 1
+% ...
+% lea r_1 1
+% store r_1 encode(i_6)
+% lea r_1 -5 //offset to first instruction
+% restrict r_1 encodePerm(e)
+
+%Activation code (i_1,...,i_6)
+% i_1 = move r_t1 pc
+% i_2 = lea r_t1 -2
+% i_3 = load r_env r_t1
+% i_4 = lea r_t1 1
+% i_5 = load r_t1 r_t1
+% i_6 = jmp r_t1
+
               \item[\texttt{load $r$ $x$}] Assumes environment capability available in register $r_{\var{env}}$.
@@
   \begin{itemize}
-  \item $c_\malloc$ satisfies the specification for malloc
+  \item $c_\malloc$ satisfies the specification for malloc with $\iota_{\malloc,0}$
   \end{itemize}
@@
           \]
-          % TODO elaborate on what happens here using anti reduction lemma and malloc spec.
-          Use Lemma~\ref{lem:anti-red-obs}, so let $\ms_f'$ be given, then
-          \[
-            (\reg_1\update{\pcreg}{\updatePcPerm{c_{g1}}},\ms_1 \uplus \ms_f') \step[i] (\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\env \uplus \ms_x \uplus \ms_\cls \uplus \ms_f')
-          \]
-          for some $i$. Where
-          \begin{itemize}
-          \item $\memSat[n_2-i]{\ms_2 \uplus \ms_\malloc'}{W_1\update{0}{\iota_\malloc}} $
-          \item $\iota_\malloc \futurewk \iota_{\malloc,0}$
+          Using Lemma~\ref{lem:malloc-works}, Lemma~\ref{lem:crtcls-works}, Lemma~\ref{lem:anti-red-obs} (and some others),
+          it suffices to show
+          \[
+            \npair[n_2']{(\reg_2,\ms_2 \uplus \ms_\malloc' \uplus \ms_\cls \uplus \ms_x)} \in \observations(W_2)
+          \]
+          where
+          \[
+            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms_\cls)}\update{i_2}{\iota_x}
+          \]
+          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is the region in Definition~\ref{def:iotax-region} which is a region by Lemma~\ref{lem:iotax-is-a-region}. Also
+          \begin{itemize}
+          \item $\iota_\malloc \futurestr \iota_\malloc'$
+          \item $c_x = ((\rwx,\glob),x,x,x)$
           \item $\ms_x = [x \mapsto 0]$
-          \item $c_x = ((\rwx,\glob),x,x,x)$
+          \item $\memSat[n_2']{\ms_2 \uplus \ms_\malloc' \uplus \ms_\cls \uplus \ms_x}{W_2}$
+          \item $c_\env = ((\rwx,\glob,\env,\env,\env))$
           \item $\ms_\env = [\env \mapsto c_x]$
-          \item $c_\env = ((\rwx,\glob,\env,\env,\env))$
-          \item $c_{f4}$ is $c_{g1}$ updated to point at $f4$.
-          \item $\ms_\cls$ contains $c_\env$, the $c_{f4}$, and the activation code.
-          \item $c_\cls$ is a global enter capability for $\ms_\cls$ pointing to the activation code.
-          \item $\reg_2 = reg_0\update{\pcreg}{\updatePcPerm{\reg_1(r_0)}}\update{r_0}{\reg_1(r_0)}\update{r_1}{c_\cls}$
-          \end{itemize}
-          For world
-          \[
-            W_2 = W_1\update{0}{\iota_\malloc}\update{i_1}{\iota^\sta (\perm,\ms_\env \uplus \ms_\cls)}\update{i_2}{\iota_x}
-          \]
-          where $i_1,i_2\not\in\dom(W_1)$ and $i_1 \neq i_2$ and $\iota_x$ is the region in Definition~\ref{def:iotax-region} which is a region by Lemma~\ref{lem:iotax-is-a-region}.
+          \item $c_{f4} = ((\rwx,\glob),\mathtt{g1} - \olf,\mathtt{4f},\mathtt{f4})$
+          \item $\ms_\cls = \ms_\env \uplus \ms_\act$
+          \item 
+            \begin{equation*}
+              \reg_2(r) =
+              \begin{cases}
+                \updatePcPerm{\reg_1(r_0)} & r = \pcreg \\
+                \reg_1(r_0) & r = r_0 \\
+                c_\cls & r = r_1 \\
+                0 & \text{otherwise}
+              \end{cases}
+            \end{equation*}
+          \end{itemize}
+          Finally assume Hyp-Act:
+            \begin{multline}
+            \label{eq:g1-hyp-act}
+            \forall \reg,\ms \ldotp \reg(\pcreg)=c_\cls \Rightarrow\\
+            \exists j \ldotp \forall \ms_f \ldotp (\reg,\ms \uplus \ms_\cls \uplus \ms_f) \step[j] (\reg\update{\pcreg}{\updatePcPerm{c_{f4}}}\update{r_\env}{c_\env},\ms \uplus \ms_\cls \uplus \ms_f)
+          \end{multline}
           Show
@@
                   \end{equation}
-                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W_3}}$ (using Lemma~\ref{lem:priv-mono-like2})
-
-                  Use Lemma~\ref{lem:anti-red-obs}, let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern.
+                  Let $\ms_3^p$ and $\ms_3^t$ be memory segments such that $\ms_3 = \ms_3^p \uplus \ms_3^t$ and $\memSat[n_4]{\ms_3^p}{\revokeTemp{W_3}}$ (using Lemma~\ref{lem:priv-mono-like2}).
+By $\memSat[n_4]{\ms_3}{W_3}$ and $W_3 \futurestr W_2$, we know $\ms_\cls \subseteq \ms_3^p$, so using Hyp-Act(\ref{eq:g1-hyp-act}), we get $j$ such that
+                  \[
+                    \forall \ms_f \ldotp (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f) \step[j] (\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t \uplus \ms_f)
+                  \]
+                  Using Lemma~\ref{lem:anti-red-obs} it suffices to show
+                  \[
+                    \npair[n_4]{(\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t)} \in \observations(W_3)
+                  \]
+                  Use Lemma~\ref{lem:anti-red-obs} again.
+This time let $\ms_f''$ be given and take $\ms_r$ to be the part of $\ms_3^t$ that $\reg_3(r_\stk)$ does not govern.
 By the operational semantics, we know\footnote{the execution may fail, but then the configuration is trivially in the observation relation.}
                   \[
-                    (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f'') \step[j] (\reg_4,\ms_4 \uplus \ms_3^t \uplus \ms_f'')
+                    (\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t \uplus \ms_f'') \step[j']
+                    (\reg_4,\ms_4 \uplus \ms_3^t \uplus \ms_f'')
                   \]
@@
                     \begin{itemize}
-                    \item \texttt{prepstack} did not fail, so the capability must be $\rwlx$ and follow the stack convention.
-                    \end{itemize}
-                  \item $\reg_3(r_1)$ is a $\glob$ capability.
+                    \item \texttt{prepstack} did not fail, so the stack capability must be $\rwlx$ and follow the stack convention.
+                    \end{itemize}
+                  \item $\reg_4(r_1)$ is a $\glob$ capability.
                     \begin{itemize}
@@
                   \item $\ms_4(x) = 0$
-                  \item $\reg_3(r_\env) = c_\env$
+                  \item $\reg_4(r_\env) = c_\env$
                   \end{itemize}
@@
                   \[
-                    \memSat[n_4-j]{\ms_4}{W_4}
+                    \memSat[n_4-j-j']{\ms_4}{W_4}
                   \]
@@
                   \[
-                    \npair[n_4-j]{(\reg_4,\ms_4 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused)} \in \observations(W_4)
+                    \npair[n_4-j-j']{(\reg_4,\ms_4 \uplus \ms_r \uplus \emptyset \uplus \ms_\unused)} \in \observations(W_4)
                   \]
@@
                   \begin{enumproof}
-                    \item $\memSat[n_4-j]{\ms_4}{\revokeTemp{W_4}}$: follows from $W_4 = \revokeTemp{W_4}$
+                    \item $\memSat[n_4-j-j']{\ms_4}{\revokeTemp{W_4}}$: follows from $W_4 = \revokeTemp{W_4}$
                     \item Hyp-Callee \label{pf:g1-first-hyp-callee}\\
@@
                       \[
-                        \npair[n_4-j]{\updatePcPerm{reg_3(r_1)}} \in \stder(W_5)
+                        \npair[n_4-j-j']{\updatePcPerm{reg_3(r_1)}} \in \stder(W_5)
                       \]
@@
                       \begin{enumproof}
-                        \item $\memSat[n_4-j-1]{\ms_5}{W_5}$ for some $\ms_5$ which is one of the assumptions of Hyp-Callee.
-                        \item $\npair[n_4-j-1]{\reg_5} \in \stdrr(W_5)$ where $\reg_5$ is as described in the \texttt{scall} lemma Hyp-callee premise.\\
+                        \item $\memSat[n_4-j-j'-1]{\ms_5}{W_5}$ for some $\ms_5$ which is one of the assumptions of Hyp-Callee.
+                        \item $\npair[n_4-j-j'-1]{\reg_5} \in \stdrr(W_5)$ where $\reg_5$ is as described in the \texttt{scall} lemma Hyp-callee premise.\\
                           Amounts to showing:\lau{too deeply nested for environments}\\
-                          1) $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.
+                          1) $\npair[n_4-j-j'-1]{\reg_3(r_1)} \in \stdvr(W_5)$, use Lemma~\ref{lem:stdvr-glob-priv-mono} with $\npair[n_4-j-1]{\reg_3(r_1)} \in \stdvr(W_3)$, the capability is global, and $W_5 \futurestr W_3$.
                           2) The protected return pointer and the stack capability are in the value relation by Hyp-callee assumptions.
@@
                       \[
-                        \npair[n_4-j-1]{(\reg_5,\ms_5)} \in \observations(W_5)
+                        \npair[n_4-j-j'-1]{(\reg_5,\ms_5)} \in \observations(W_5)
                       \]
@@
                       \begin{itemize}
-                      \item $n_5 \leq n_4-j-2$
+                      \item $n_5 \leq n_4-j-j'-2$
                       \item $W_6 \futurewk \revokeTemp{W_4}$
@@
 \begin{lemma}[$\mathtt{crtcls}$ works]
+  \label{lem:crtcls-works}
   If
@@
     \item $\ms_\cls = \ms_\act \uplus \ms_\env$
+    \item $c_\cls = ((\entry,\glob),\dots)$
     \item $c_\env = ((\readwrite,\glob),\start_\env,\addrend_\env,\start_\env)$

### 2017-02-23 local-capability-machine.tex
added lemma about rwlx and pwl regions and used it in g1 proof

@@
               \end{itemize}
-              we get that there exists a region, $r_{\adv \stk}$ in $W_3$ such that
-              \[
-                W_3(r_{\adv \stk}) = \iota_{\stk_a,\stk_b}^\pwl
+              using Lemma~\ref{lem:pwl-stack} we get that there exists a region, $r_{\adv \stk}$ such that
+              \[
+                W_3(r_{\adv \stk}) \nequal \iota_{\stk_a,\stk_b}^\pwl
               \]
@@
 \begin{lemma}
+  \label{lem:pwl-stack}
+  For all $W$, $\start$, $\addrend$, $n$, $\ms$ if
+  \begin{itemize}
+  \item $\memSat{\ms}{W}$
+  \item $\npair{((\rwlx,\gl),\start,\addrend,\addr)} \in \stdvr(W)$
+  \item $\start \leq \addrend$
+  \end{itemize}
+  then
+  \[
+    \exists r, \start',\addrend' \ldotp [\start,\addrend] \subseteq [\start',\addrend'] \land W(\var{r}) \nequal \iota^\pwl_{\start',\addrend'}
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:pwl-stack}]
+  Assume
+  \begin{enumproof}
+    \item $\npair{((\rwlx, \gl),b,e,a)} \in \stdvr(W)$ \label{ass:pwl-stack:val-rel}
+    \item $\memSat{\ms}{W}$ \label{ass:pwl-stack:mem-sat}
+  \end{enumproof}
+  From Assumption~\ref{ass:pwl-stack:val-rel}, we get $r_1$, $r_2$, $b_1$, $b_2$, $e_1$ and $e_2$ such that
+  \begin{enumproof}[resume]
+    \item $r_1 \in \localityReg(\gl,W)$
+    \item $r_2 \in \localityReg(\gl,W)$
+    \item $[b,e] \subseteq [b_1,e_1]$\label{ass:pwl-stack:b1}
+    \item $[b,e] \subseteq [b_2,e_2]$ \label{ass:pwl-stack:b2}
+    \item $W(r_1) \nsubsim \iota^\pwl_{b_1,e_1}$ \label{ass:pwl-stack:subsim}
+    \item $W(r_2) \nsupsim \iota^\pwl_{b_2,e_2}$ \label{ass:pwl-stack:supsim}
+    \item $W(r_2)$ is address-stratified. \label{ass:pwl-stack:add-stra}
+  \end{enumproof}
+From Assumption~\ref{ass:pwl-stack:mem-sat}, we get partitionen $P$ s.t.
+\[
+  \memSat[n,p]{\ms}{W}
+\]
+Say $P(r_1) = \ms_1$ and $P(r_2) = \ms_2$.
+First from $\npair{\ms_1} \in W(r_1).H\;W(r_1).s\;\xi^{(-1)}(W)$ using \label{ass:pwl-stack:subsim}, we get $\npair{\ms_1} \in H^\pwl_{b_1,e_1}\;1\;\xi^{(-1)}(W)$ which means $\dom(\ms_1) = [b_1,e_1]$.
+
+Second we know $\npair{[b_2 \mapsto 0, \dots , e_2 \mapsto 0]} \in H^\pwl_{b_2,e_2}\;1\;\xi^{(-1)}(W)$ and $\npair{\ms_2} \in W(r_2).H\;W(r_2).s\;\xi^{(-1)}(W)$ which by Assumption~\ref{ass:pwl-stack:supsim} and \ref{ass:pwl-stack:add-stra} means $\dom(\ms_2) = [b_2,e_2]$.
+
+Now assume for contradition $r_1 \neq r_2$, then we have a contradiction with $\memSat[n,p]{\ms}{W}$ because $\ms_1$ and $\ms_2$ are not disjoint (by Assumptions~\ref{ass:pwl-stack:b1} and \ref{ass:pwl-stack:b2}).
+So $r_1=r_2$ which also means $[b_1,e_1] = [b_2,e_2]$, so from Assumption~\ref{ass:pwl-stack:subsim} and \ref{ass:pwl-stack:supsim}, we get $W(r_1) \nsim \iota^\pwl_{b_1,e_1}$ which by Lemma~\ref{lem:pwl-nsim-view} means $W(r_1) \nequal \iota^\pwl_{b_1,e_1}$
+\end{proof}
+
+\begin{lemma}
   \label{lem:hpwl-mono}
@@
   \begin{equation*}
-    \revokeTemp(W)(r) \text{ is address-stratified}
+    \revokeTemp{W}(r) \text{ is address-stratified}
   \end{equation*}

### 2017-02-23 local-capability-machine.tex
Added another sanity lemma
fixed ref issues (by reinserting small lemma in proof that was removed by mistake)

@@
           \item $\npair[n-i]{c_\malloc} \in \stdvr(W)$: Follows from Lemma~\ref{lem:malloc-in-vr}.
-          \item $\npair[n-i]{c_\adv} \in \stdvr(W)$: Follows from Theorem~\ref{thm:ftlr} and Lemma~\ref{lem:nwlp-subset-pwl}.
+          \item $\npair[n-i]{c_\adv} \in \stdvr(W)$: \\
+            We will show
+            \begin{equation}
+              \label{eq:f3:adv-in-stder}
+              \forall W' \futurestr W \ldotp \npair{c_\adv} \in \stdvr(W')
+            \end{equation}
+            which will give us what we need using downwards closure as well as a result for later use.
+
+            Let $W' \futurestr W$ be given and show
+            \[
+              \npair{(\start_\adv,\addrend_\adv,\start_\adv+\olf)} \in \entryCond{}(\glob)(W')
+            \]
+            to this end let $W'' \futurestr W'$ and $n' < n$ be given and show
+            \[
+              \npair[n']{\updatePcPerm{c_\adv}} \in \stder(W'')
+            \]
+            This follows from the FTLR (Theorem~\ref{thm:ftlr}) if we can show
+            \[
+              \npair[n']{\start_\adv,\addrend_\adv} \in \readCond{}(\glob)(W'')
+            \]
+            $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}$ governs the adversary, so the result follows from Lemma~\ref{lem:nwlp-subset-pwl}.
           \end{enumproof}
@@
   \item $\memSat{\ms}{W}$
-  \item $\npair{((\rwlx,\gl),\start,\addrend,\addr)} \in \stdvr(W)$
+  \item $\npair{\stdcap} \in \stdvr(W)$
   \item $\start \leq \addrend$
+  \item $\perm \in \{ \rwlx, \rwx \}$
   \end{itemize}
@@
 which follows from \ref{pf:heap-sat:interp}, $n' \leq n$, and $H(s)(\xi^{-1}(W))$ is a $\UPred{\MemSegments}$.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:wl-local}
+  If
+  \begin{itemize}
+  \item $\memSat{\ms}{W}$
+  \item $\npair{\stdcap} \in stdvr(W)$
+  \item $\start \leq \addrend$
+  \item $\perm \in \{\rwlx,\rwl\}$
+  \end{itemize}
+  then 
+  \[
+    \gl = \local
+  \]
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:wl-local}]
+It follows as a consequence of Lemma~\ref{lem:pwl-stack}. The $n$-equality forces the region to be $\temp$, so for the region name to be in $\localityReg(\gl,W)$, the locality must be $\local$.
 \end{proof}

### 2017-03-15 csf-2017/paper.tex
inserted authors and ack., so it is ready for distribution
typos

@@
 
-% \author{%
+\author{%
   % \IEEEauthorblockN{Lau~Skorstengaard}
@@
   % \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
-% Email: birkedal@cs.au.dk}}
-\author{}
+  %   Email: birkedal@cs.au.dk}
+}
 \maketitle
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine the adversary stores his stack pointer, invokes trusted
+pointers. Imagine the adversary stores their stack pointer, invokes trusted
 code that uses part of the stack to store private data and then invokes the
@@
 the private data. If the adversary had a way of keeping hold of their old stack
-pointer, it could access the private data stored there by the trusted code, and
+pointer, it could access the private data stored there by the trusted code and
 break local-state encapsulation.
@@
 make sure the stack capability gives $\rwlx$ authority, since the system wide assumption then
-tells us that the adversary cannot have a global capabilities to our local stack.
+tells us that the adversary cannot have global capabilities to our local stack.
 
@@
 
-\begin{comment}
 \section*{Acknowledgements}
@@
 Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
-\end{comment}
 

### 2017-03-15 local-capability-machine.tex
changed the standard region to take set instead of range and defined the old one in terms of the new
typos and small corrections

@@
 \end{align*}
-Further, assume an inverse function, $\decodePermPair{}$, that decodes permissions
+Further, assume a left inverse function, $\decodePermPair{}$, that decodes permissions
 \[
@@
   \sem{\fail}(\Phi)                        & = \failed \\
-  \sem{\halt}(\Phi)                        & = (\halted,\Phi) \\
+  \sem{\halt}(\Phi)                        & = (\halted,\memheap) \\
   \sem{\jmp{\lv}}(\Phi)                    & = \updateReg{\pcreg}{\updatePcPerm{\memreg(\lv)}} \\
@@
                                              \end{cases} \\
-  \sem{\load{\refreg{r_1}}{\refheap{r_2}}}  & = 
+  \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi)  & = 
                                               \begin{cases}
@@
                                               \end{cases}\\
-  \sem{\store{\refheap{r_1}}{\refreg{r_2}}} & = 
+  \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) & = 
                                               \begin{cases}
@@
                                               \end{cases}\\
-  \sem{\move{\refreg{r_1}}{\rv}}            & = \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}}
+  \sem{\move{\refreg{r_1}}{\rv}}(\Phi)            & = \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}}
   \\
-  \sem{\lea{\refreg{r_1}}{\rv}}            & =
+  \sem{\lea{\refreg{r_1}}{\rv}}(\Phi)            & =
                                              \begin{cases}
@@
   % In the M-Machine, lea checks whether the pointer stay within the allowed range in the pointer.
-  \sem{\restricttwo{\refreg{r}}{\rv}}           & =
+  \sem{\restricttwo{\refreg{r}}{\rv}}(\Phi)           & =
                                                   \begin{cases}
@@
 \begin{align*}
-  \sem{\plus{\refreg{r_1}}{\rv_1}{\rv_2}}               & =
+  \sem{\plus{\refreg{r_1}}{\rv_1}{\rv_2}}(\Phi)               & =
                                                           \begin{cases}
@@
                                                           \end{cases}\\
-  \sem{\minus{\refreg{r_1}}{\rv_1}{\rv_2}}               & =
+  \sem{\minus{\refreg{r_1}}{\rv_1}{\rv_2}}(\Phi)               & =
                                                           \begin{cases}
@@
                                                           \end{cases}\\
-  \sem{\subseg{\refreg{r}}{\rv_1}{\rv_2}} & = 
+  \sem{\subseg{\refreg{r}}{\rv_1}{\rv_2}}(\Phi) & = 
                                             \begin{cases}
@@
   \\
-  \sem{\geta{\refreg{r_1}}{\refreg{r_2}}} & = 
+  \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) & = 
                                             \begin{cases}
@@
   \\
-  \sem{\getb{\refreg{r_1}}{\refreg{r_2}}} & = 
+  \sem{\getb{\refreg{r_1}}{\refreg{r_2}}}(\Phi) & = 
                                             \begin{cases}
@@
   \\
-  \sem{\gete{\refreg{r_1}}{\refreg{r_2}}} & = 
+  \sem{\gete{\refreg{r_1}}{\refreg{r_2}}}(\Phi) & = 
                                             \begin{cases}
@@
   \\
-  \sem{\getp{\refreg{r_1}}{\refreg{r_2}}} & = 
+  \sem{\getp{\refreg{r_1}}{\refreg{r_2}}}(\Phi) & = 
                                             \begin{cases}
@@
   \\
-  \sem{\getl{\refreg{r_1}}{\refreg{r_2}}} & = 
+  \sem{\getl{\refreg{r_1}}{\refreg{r_2}}}(\Phi) & = 
                                             \begin{cases}
@@
   \\
-  \sem{\isptr{\refreg{r}}{\rv}} & =  
+  \sem{\isptr{\refreg{r}}{\rv}}(\Phi) & =  
                                   \begin{cases}
@@
 
-              \begin{specification}[Malloc v.2]
+              \begin{specification}[Malloc Specification]
                 \label{spec:malloc}
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^\pwl : \Regions\\
-  &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
+  &\iota^\pwl : \cal{P} \fun \Regions\\
+  &\iota^\pwl\;A \defeq (\temp,1,=,=,H^\pwl \; S) \\
   \\
-  &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^\pwl : \cal{P}(\Addrs) \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
+  &H^\pwl\; A \; s \; \hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+      &\dom(\hs) = A \land \\
+      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
     \end{aligned}
@@
 \end{align*}
+Further define
+\[
+  \iota^\pwl_{\start,\addrend} \defeq \iota^\pwl([\start,\addrend])
+\]
+
 \newcommand{\wrev}[1]{\revokeTemp{#1}}
@@
 \begin{align*}
-  &\iota_{\start,\addrend}^{\nwl} : \Regions \\
-  &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
+  &\iota^\nwl : \cal{P}(\Addrs) \fun \Regions \\
+  &\iota^\nwl\; A \defeq (\temp,1,=,=,H^\nwl\;A) \\
   \\
-  &\iota_{\start,\addrend}^{\nwl,p} : \Regions \\
-  &\iota_{\start,\addrend}^{\nwl,p} \defeq (\perma,1,=,=,H^\nwl_{start,\addrend})\\
+  &\iota^{\nwl,p} : \cal{P}(\Addrs) \fun \Regions \\
+  &\iota^{\nwl,p}\;A \defeq (\perma,1,=,=,H^\nwl\;A)\\
   \\
-  &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
+  &H^\nwl : \cal{P}(\Addrs) \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
+  &H^\nwl\;A \; s \;\hat{W} \defeq \left\{\npair{\hs} \middle|
     \begin{aligned}
-      &\dom(\hs) = [\start,\addrend] \land \\
-      &\forall \addr \in [\start,\addrend] \ldotp \\
+      &\dom(\hs) = A \land \\
+      &\forall \addr \in A \ldotp \\
       &\quad \nonlocal{\ms(\addr)} \land\\ 
@@
         \right\} \union \{\npair[0]{\ms} \}
+\end{align*}
+
+Further define
+
+\begin{align*}
+  \iota^\nwl_{\start,\addrend} & \defeq \iota^\nwl([\start,\addrend])\\
+  \iota^{\nwl,p}_{\start,\addrend} & \defeq \iota^{\nwl,p}([\start,\addrend])\\
 \end{align*}
@@
 \subsubsection{LR Sanity lemmas}
+\begin{lemma}
+  \label{lem:mem-sat-ne}
+  \begin{align*}
+    &\forall \hs, n, W \nequal W' \ldotp \\
+    &\quad \heapSat[\hs]{n}{W} \land W \nequal W' \Rightarrow \heapSat[\hs]{n}{W'}
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:mem-sat-ne}]
+  %TODO
+\end{proof}
 
@@
 
+%This lemma seems weird given the previous lemmas.
 \begin{lemma}[Revoke temporary memory with stack]
@@
     & \qquad \exists \ms', \ms_r \ldotp \\
-    & \qquad \quad \memSat{\ms }{\revokeTemp{W}} \land \ms = \ms' \uplus \ms_r
-  \end{align*}
-\end{lemma}
-\begin{proof}[lem:revoke-temp-stack]
+    & \qquad \quad \memSat{\ms'}{\revokeTemp{W}} \land \ms = \ms' \uplus \ms_r
+  \end{align*}
+\end{lemma}
+\begin{proof}[Proof of Lemma~\ref{lem:revoke-temp-stack}]
   % TODO
@@
     & \quad \ms'' = \ms \uplus \ms' \land W'' = W \uplus W' \land \memSat{\ms}{W} \land \memSat{\ms'}{W'} \Rightarrow \\
-    & \qquad \heapSat[\ms \uplus \ms' ]{n}{W \uplus W'}
+    & \qquad \heapSat[\ms'']{n}{W''}
   \end{align*}
@@
   \[
-    \memSat{\ms}{[i \mapsto \iota^\pwl(\dom(\ms))]}
+    \memSat{\ms}{[i \mapsto \iota(\dom(\ms))]}
   \]

### 2017-03-20 local-capability-machine/capability-machine.tex
restructured folder

--- 2016-11-18 capability-machine.tex
@@
 Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
-\lars{So this means that the value returned by *l is 1, correct ? We
-  could add an assert statement to the example to show it more directly.}
-\lau{Yes, and done.}
-
-\lars{What about an even simpler example, such as the one below:}
+
 \begin{verbatim}
@@
 \end{verbatim}
-\lau{This example should certainly also hold true work, but it does not need local capabilities (it might serve well as a sanity check). }
+
+Correctness lemma not assuming stack
+\begin{lemma}
+  For all $n$,
+  let
+  \begin{align*}
+    c_\adv & = ((\entry,\local),\start_\adv,\addrend_\adv,\start_\adv) \\
+    c_f & = ((\exec,\local),\start_f,\addrend_f,\start_f)
+  \end{align*}
+  if
+  \[
+    (\reg_0\update{\pcreg}{c_f}\update{r_1}{c_\adv},\hs_\adv \uplus \hs_f \uplus \hs_\malloc \uplus h_\flag \uplus h_\var{frame}) \step[n] (\halted, h')
+  \]
+  where
+\lau{TODO: finish this}
+\end{lemma}
 
@@
 \newcommand{\rio}{r_{io}}
-\newcommand{\adv}{\codelabel{adv}}
+\renewcommand{\adv}{\codelabel{adv}}
 \newcommand{\advb}{\var{adv_{base}}}

### 2017-03-20 fully-abstract-comp/technical-report.tex
Technical report added. Typeset some of the blackboard notes.

\documentclass[a4paper]{article}
\usepackage[a4paper]{geometry}
\usepackage[utf8]{inputenc}

% Math packages
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{array}

% Command control packages
\usepackage{ifthen}
\usepackage{ifpdf}

% Listings
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

% Tikz
\usepackage{tikz}

%%% Function arrows
\newcommand{\fun}{\rightarrow}

%%% Instruction formatting
\newcommand{\sourcecolortext}{blue}
\newcommand{\sourcecolor}[1]{\color{blue}}
\newcommand{\src}[1]{{\sourcecolor{} #1}}
\newcommand{\targetcolortext}{green}
\newcommand{\targetcolor}[1]{\color{green}}
\newcommand{\trg}[1]{{\targetcolor{} #1}}

\newcommand{\zinstr}[1]{#1}
\newcommand{\oneinstr}[2]{
  \ifthenelse{\equal{#2}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2}
}
\newcommand{\twoinstr}[3]{
  \ifthenelse{\equal{#2#3}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2 \; #3}
}
\newcommand{\threeinstr}[4]{
  \ifthenelse{\equal{#2#3#4}{}}
  {\zinstr{#1}}
  {\zinstr{#1} \; #2 \; #3 \; #4}
}

%%% Source language
% No arguments
\newcommand{\sfail}{\zinstr{\src{fail}}}
\newcommand{\shalt}{\zinstr{\src{halt}}}
\newcommand{\sreturn}{\zinstr{\src{return}}}

% One argument
\newcommand{\sjmp}[1]{\oneinstr{\src{jmp}}{#1}}
\newcommand{\spush}[1]{\oneinstr{\src{push}}{#1}}
\newcommand{\spop}[1]{\oneinstr{\src{pop}}{#1}}
\newcommand{\ssload}[1]{\oneinstr{\src{sload}}{#1}}

% Two arguments
\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
\newcommand{\sisptr}[2]{\twoinstr{\src{isptr}}{#1}{#2}}
\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
\newcommand{\sgetl}[2]{\twoinstr{\src{getl}}{#1}{#2}}
\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}

% Three arguments
\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
\newcommand{\ssubseg}[3]{\threeinstr{\src{subseg}}{#1}{#2}{#3}}
\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}

%%% Target language
% No arguments
\newcommand{\tfail}{\zinstr{\trg{fail}}}
\newcommand{\thalt}{\zinstr{\trg{halt}}}

% One argument
\newcommand{\tjmp}[1]{\oneinstr{\trg{jmp}}{#1}}

% Two arguments
\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
\newcommand{\tisptr}[2]{\twoinstr{\trg{isptr}}{#1}{#2}}
\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
\newcommand{\tgetl}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}

% Three arguments
\newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
\newcommand{\tsubseg}[3]{\threeinstr{\trg{subseg}}{#1}{#2}{#3}}
\newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
\newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}

%%% Domains
\newcommand{\plaindom}[1]{\mathrm{#1}}
\newcommand{\RegName}{\plaindom{RegisterName}}

\newcommand{\nats}{\mathbb{N}}

%%% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\rn}{\var{rn}}
\newcommand{\stk}{\var{stk}}

%%% Named registers
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\rstk}{\mathrm{r}_\mathrm{stk}}
\newcommand{\rO}{\mathrm{r}_0}

%%% Permissions
\newcommand{\plainperm}[1]{\mathrm{#1}}
\newcommand{\rwlxo}{\plainperm{rwlxo}}
\newcommand{\rwlo}{\plainperm{rwlo}}
\newcommand{\rwl}{\plainperm{rwl}}
\newcommand{\rwxo}{\plainperm{rwxo}}
\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
\newcommand{\rxo}{\plainperm{rxo}}
\newcommand{\ro}{\plainperm{ro}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\noperm}{\plainperm{0}}
\newcommand{\nopermo}{\plainperm{0o}}
\newcommand{\enter}{\plainperm{e}}

%%% Braces
\newcommand{\comp}[1]{[#1]}

%%% Functions

\begin{document}
\section{The two capability machines}
\subsection{Domains}
\[
  \begin{array}{rcl}
    \{\pcreg,\rO,\rstk \} & \subseteq & \RegName
  \end{array}
\]

\subsection{Syntax}
The target machine is a simple capability machine with memory capabilities, local capabilities and enter capabilities. The syntax of the instructions of the target machine is defined as follows:
\[
\begin{array}{rcl}
n &\in & \nats \\
\trg{r} &\in &  \RegName  \src{\setminus \{\rO,\rstk\}}\\
\trg{\rn} &::= &  \trg{r} \mid n \\
\trg{i} &::= & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}}
\end{array}
\]
Where the \sourcecolortext{} only applies for the source language which we define below.

The source machine is also a capability machine with memory capabilities, local capabilities and enter capabilities. Unlike the target machine, the source machine is going to have a built in stack. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rO,\rstk\}\\
    \src{\rn} &::= & \src{r} \mid n \\
    \src{i} & ::= &  \trg{i} \mid \spop{\src{r}} \mid \spush{\src{\rn}} \mid \ssload{\src{\rn}} \mid \scall{\src{r}}{\src{\rn}} \mid \sreturn
  \end{array}
\]
There are two syntactic differences between the source language and the target language: (1) The source language does not permit register names $\rO$ and $\rstk$ to be used because they will play a special role in the compilation scheme. (2) The source language has explicit call/return instructions as well as the stack manipulation instructions $\spop{}$, $\spush{}$, $\ssload{}$, and $\scall{}{}$. We will define the semantic of these below.

\subsection{Permissions}

\begin{figure}[!h]
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (7) {$\rwlxo$};
    \node[main node] (1) [below of=7] {$\rwxo$};
    \node[main node] (2) [below of=1] {$\rxo$};
    \node[main node] (3) [below of=2] {$\enter$};

    \node[main node] (9) [below left of=7] {$\rwlo$};
    \node[main node] (10) [below of=9] {$\rwo$};
    \node[main node] (11) [below of=10] {$\ro$};
    \node[main node] (12) [below of=11] {$\nopermo$};

    \node[main node] (8) [above left of=9] {$\rwl$};
    \node[main node] (4) [below of=8] {$\rw$};
    \node[main node] (5) [below of=4] {$\readonly$};
    \node[main node] (6) [below of=5] {$\noperm$};

    \path[every node/.style={font=\sffamily\small}]
    (7) edge (1)
    (8) edge (4)
    (1) edge (2)
    (2) edge (3)
    (4) edge (5)
    (5) edge (6)

    (9) edge (10)
    (10) edge (11)
    (11) edge (12)

    (9) edge (7)
    (10) edge (1)
    (11) edge (2)
    (12) edge (3)

    (9) edge (8)
    (10) edge (4)
    (11) edge (5)
    (12) edge (6);
  \end{tikzpicture}

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}

\subsection{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]
\clearpage
\section{Examples}
\subsection{Capabilitiy opacity}
The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler corretness
\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
p1 ::= if r1 is length 2 then
         call r1 with the following callback in r5:
           {put $\textit{\texttt{diverging}}$ closure in r2;
            return};
         halt
       else diverge

p2 ::= if r1 is length 2 then
         call r1 with the following callback in r5:
           {put $\textit{\texttt{terminating}}$ closure in r2;
            return};
         halt
       else diverge
\end{lstlisting}
The diverging closure could just contain $\sjmp{\pcreg}$, which diverges. The terminating closure could just be $\sreturn$.

The context with $r1$ as an executable capability pointing to:
\begin{lstlisting}
$\scall{r5}{0}$
$\sjmp{r5}$
\end{lstlisting}
distinguishes the two contexts, but two instructions are not enough to do the same at the target level. We would not have enough instructions to set up a proper return pointer for the compiled return to use.

\clearpage
\section{Back translation}
We want to do the back translation by interpreting the programs.
\end{document}

### 2017-03-21 fully-abstract-comp/technical-report.tex
spelling

@@
 \newcommand{\rstk}{\mathrm{r}_\mathrm{stk}}
-\newcommand{\rO}{\mathrm{r}_0}
+\newcommand{\rO}{\mathrm{r}_\mathrm{ret}}
 
@@
 \section{Examples}
-\subsection{Capabilitiy opacity}
-The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler corretness
+\subsection{Capability opacity}
+The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
 \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 

### 2017-03-21 local-capability-machine/local-capability-machine.tex
Kyndylan mail corrections

--- 2017-03-15 local-capability-machine.tex
@@
 \usepackage{mathpartir}
+\usepackage{hyperref}
 \usepackage{listings}
-\usepackage{hyperref}
 \lstset{
@@
   \end{tikzpicture}
-  \caption{Permission hierarchy}
+  \caption{Locality hierarchy}
   \label{fig:glob-hier}
@@
 
+\begin{comment}
 After adding local capabilities, most of the definitions have changed slightly to take into account the new form the capabilities take.
@@
 Whether a $\mathtt{restrict}$ succeeds depends on the ordering, so the real change lies in the new ordering, which is described further above (it is the point-wise ordering of the pair).
-
+\end{comment}
 Define the following macros: $\mathtt{restrict}$, $\mathtt{subseg}$, and $\mathtt{lea}$ that does not overwrite the source register.
@@
   If $\Phi \step[n] \Phi'$ and $\Phi \step[n] \Phi''$, then $\Phi' = \Phi''$.
-  If $\Phi \step[n] \Phi'$ and $\Phi \step[n'] (\halted,\mem'')$, then $n \geq n'$
+  If $\Phi \step[n] \Phi'$ and $\Phi \step[n'] (\halted,\mem'')$, then $n \leq n'$
   and $\Phi' \step[n'-n] (\halted,\mem'')$.
@@
   \iota^\cnst (v,n) &= (v,1,=,=,H^\cnst\;n)\\
-  H^\cnst \; n \; s \; \hat{W} = &  \{\npair{\ms} \mid n > 0 \land \forall \addr \in \dom(\ms) \ldotp \ms(\addr) = n \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
+  H^\cnst \; n' \; s \; \hat{W} = &  \{\npair{\ms} \mid n > 0 \land \forall \addr \in \dom(\ms) \ldotp \ms(\addr) = n' \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
 \end{align*}

### 2017-03-22 fully-abstract-comp/technical-report.tex
more stuff

@@
 \usepackage{array}
+\usepackage{stmaryrd}
 
@@
 
-%%% Function arrows
+%%% Comments
+% Comments
+\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
+\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
+\newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+
+%%% Math notation
+\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
+\newcommand{\defbnf}{::=}
+\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
+\newcommand{\ssem}[2][\Phi]{\sem{#2}_{\mathrm{src}}(#1)}
+
+% Function arrows
 \newcommand{\fun}{\rightarrow}
+\newcommand{\parfun}{\rightharpoonup}
 
@@
 \newcommand{\sjmp}[1]{\oneinstr{\src{jmp}}{#1}}
+\newcommand{\ssetatob}[1]{\oneinstr{\src{seta2b}}{#1}}
 \newcommand{\spush}[1]{\oneinstr{\src{push}}{#1}}
 \newcommand{\spop}[1]{\oneinstr{\src{pop}}{#1}}
-\newcommand{\ssload}[1]{\oneinstr{\src{sload}}{#1}}
 
@@
 \newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
+\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
 \newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
@@
 \newcommand{\tjmp}[1]{\oneinstr{\trg{jmp}}{#1}}
+\newcommand{\tsetatob}[1]{\oneinstr{\trg{seta2b}}{#1}}
 
@@
 \newcommand{\plaindom}[1]{\mathrm{#1}}
-\newcommand{\RegName}{\plaindom{RegisterName}}
 
 \newcommand{\nats}{\mathbb{N}}
+\newcommand{\ints}{\mathbb{Z}}
+
+%%% Updates
+\newcommand{\update}[2]{[#1 \mapsto #2]}
+\newcommand{\updReg}[3][\Phi]{#1\update{\sreg.#2}{#3}}
+\newcommand{\updPc}[2][\Phi]{\updReg[#1]{\pcreg}{#2}}
+
+%%% Source dom
+\newcommand{\sourcedom}[1]{\mathrm{#1}}
+\newcommand{\sRegName}{\sourcedom{RegisterName}}
+\newcommand{\sAddr}{\sourcedom{Addr}}
+\newcommand{\sPerm}{\sourcedom{Perm}}
+\newcommand{\sCap}{\sourcedom{Cap}}
+\newcommand{\sWord}{\sourcedom{Word}}
+\newcommand{\sMem}{\sourcedom{Memory}}
+\newcommand{\sReg}{\sourcedom{RegisterFile}}
+\newcommand{\sStk}{\sourcedom{Stack}}
+\newcommand{\sConf}{\sourcedom{Conf}}
+\newcommand{\sExecConf}{\sourcedom{ExecConf}}
+\newcommand{\sGlobal}{\sourcedom{Global}}
+\newcommand{\sMemSeg}{\sourcedom{MemorySegment}}
+\newcommand{\sStkFrame}{\sourcedom{StackFrame}}
+\newcommand{\sStack}{\sourcedom{Stack}}
+
+\newcommand{\smem}{\sourcedom{mem}}
+\newcommand{\sreg}{\sourcedom{reg}}
+\newcommand{\sstk}{\sourcedom{stk}}
+\newcommand{\smsstk}{\sourcedom{ms_{stk}}}
+\newcommand{\sstkframe}{\sourcedom{frame}}
+\newcommand{\sopc}{\sourcedom{opc}}
+\newcommand{\sastk}{\sourcedom{a_{stk}}}
+\newcommand{\sperm}{\sourcedom{perm}}
+\newcommand{\sgl}{\sourcedom{g}}
+\newcommand{\sbase}{\sourcedom{base}}
+\newcommand{\send}{\sourcedom{end}}
+\newcommand{\saddr}{\sourcedom{a}}
+\newcommand{\scap}{\sourcedom{c}}
+\newcommand{\sms}{\sourcedom{ms}}
+
+\newcommand{\stkptr}[4]{\sourcedom{stack\text{-}ptr}(#1,#2,#3,#4)}
+\newcommand{\retptr}{\sourcedom{ret\text{-}ptr}}
+
+\newcommand{\failed}{\mathrm{failed}}
+\newcommand{\halted}{\mathrm{halted}}
+
+%%% Target domain
+\newcommand{\targetdom}[1]{\mathrm{#1}}
+\newcommand{\tRegName}{\targetdom{RegisterName}}
 
@@
 \newcommand{\rn}{\var{rn}}
+\newcommand{\reg}{\var{reg}}
+\newcommand{\mem}{\var{mem}}
+\newcommand{\ms}{\var{ms}}
+\newcommand{\pc}{\var{pc}}
 \newcommand{\stk}{\var{stk}}
+\newcommand{\opc}{\var{opc}}
 
@@
 
+%%% locality
+\newcommand{\plainlocality}[1]{\mathrm{#1}}
+\newcommand{\glob}{\plainlocality{global}}
+\newcommand{\local}{\plainlocality{local}}
+
 %%% Permissions
 \newcommand{\plainperm}[1]{\mathrm{#1}}
+\newcommand{\rwlx}{\plainperm{rwlx}}
+\newcommand{\rwx}{\plainperm{rwx}}
+\newcommand{\rx}{\plainperm{rx}}
 \newcommand{\rwlxo}{\plainperm{rwlxo}}
@@
 \newcommand{\enter}{\plainperm{e}}
+\newcommand{\entero}{\plainperm{eo}}
 
@@
 %%% Functions
+\newcommand{\plainfun}[2]{
+  \ifthenelse{\equal{#2}{}}
+  {\mathit{#1}}
+  {\mathit{#1}(#2)}
+}
+\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
+\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
+\newcommand{\clearms}[1]{\plainfun{clear_{ms}}{#1}}
+\newcommand{\clearlocreg}[1]{\plainfun{clear_{loc,reg}}{#1}}
 
 \begin{document}
+\section{TODO}
+\begin{itemize}
+\item Linking
+\item Component
+\item Cross-language LR
+\end{itemize}
+
 \section{The two capability machines}
 \subsection{Domains}
+Source language domains:
+\[
+  \begin{array}{rrcl}
+   \saddr,\sbase \in & \sAddr & \defeq & \nats \\
+    &\sWord & \defeq & \ints \uplus \sCap\\
+   \sperm \in& \sPerm & \defbnf & \dots \\
+    \sgl \in&\sGlobal & \defbnf & \glob \mid \local \\
+    &\sCap &::= & \left\{((\sperm,\sgl),\sbase,\send,\saddr) \mid \send \in \sAddr \cup \{\infty\}\right\} {\sourcecolor{} \uplus}\\
+    & & & {\sourcecolor{} \{ \stkptr{\sperm}{\sbase}{\send}{\saddr} \}\uplus}\\ 
+    & & & {\sourcecolor{} \{ \retptr \}}\\ 
+    &\sRegName & \defbnf &\pcreg \mid \rO \mid \rstk \mid \dots \\
+    &\sReg & \defeq & \sRegName \fun \sWord\\
+    & \sMem & \defeq & \sAddr \fun \sWord \\
+    & \sMemSeg & \defeq & \sAddr \parfun \sWord \\
+    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \sStkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \sCap \times \sMemSeg \times \sAddr}\\
+    {\sourcecolor{} \sstk \in}& {\sourcecolor{} \sStack} & {\sourcecolor{} \defeq} & {\sourcecolor{} \sStkFrame^*} \\
+    \Phi \in & \sExecConf & \defeq & \sMem \times \sReg {\sourcecolor{} \; \times \; \sStack \times \sMemSeg }\\
+    &\sConf & \defeq & \sExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \sMem)
+  \end{array}
+\]
+The target language domains are all the non blue parts in the above.
+
+More convenient definitions
 \[
   \begin{array}{rcl}
-    \{\pcreg,\rO,\rstk \} & \subseteq & \RegName
+    \Phi(r) & \defeq & \Phi.\sreg(r)
   \end{array}
 \]
+where $r\in \sRegName$
 
 \subsection{Syntax}
-The target machine is a simple capability machine with memory capabilities, local capabilities and enter capabilities. The syntax of the instructions of the target machine is defined as follows:
+The target machine is[dashed] a simple capability machine with memory capabilities, local capabilities and enter capabilities.
+The syntax of the instructions of the target machine is defined as follows:
 \[
@@
 n &\in & \nats \\
-\trg{r} &\in &  \RegName  \src{\setminus \{\rO,\rstk\}}\\
+\trg{r} &\in &  \tRegName \\
 \trg{\rn} &::= &  \trg{r} \mid n \\
@@
       & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
-      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}}
+      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}
 \end{array}
 \]
-Where the \sourcecolortext{} only applies for the source language which we define below.
 
@@
     n & \in & \nats \\
-    \src{r} &\in &  \RegName \setminus \{\rO,\rstk\}\\
+    \src{r} &\in &  \sRegName \setminus \{\rO,\rstk\}\\
     \src{\rn} &::= & \src{r} \mid n \\
-    \src{i} & ::= &  \trg{i} \mid \spop{\src{r}} \mid \spush{\src{\rn}} \mid \ssload{\src{\rn}} \mid \scall{\src{r}}{\src{\rn}} \mid \sreturn
+    \src{i} & ::= &  \trg{i} \mid \scall{\src{r}}{\src{\rn}}
   \end{array}
 \]
-There are two syntactic differences between the source language and the target language: (1) The source language does not permit register names $\rO$ and $\rstk$ to be used because they will play a special role in the compilation scheme.
-(2) The source language has explicit call/return instructions as well as the stack manipulation instructions $\spop{}$, $\spush{}$, $\ssload{}$, and $\scall{}{}$.
-We will define the semantic of these below.
-
+There is one suntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
 \subsection{Permissions}
-
 \begin{figure}[!h]
@@
   \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (7) {$\rwlxo$};
-    \node[main node] (1) [below of=7] {$\rwxo$};
-    \node[main node] (2) [below of=1] {$\rxo$};
-    \node[main node] (3) [below of=2] {$\enter$};
+    \node[main node] (rwlx) {$\rwlx$};
+    \node[main node] (rwx) [below of=rwlx] {$\rwx$};
+    \node[main node] (rx) [below of=rwx] {$\rx$};
+    \node[main node] (e) [below of=rx] {$\enter$};
 
-    \node[main node] (9) [below left of=7] {$\rwlo$};
-    \node[main node] (10) [below of=9] {$\rwo$};
-    \node[main node] (11) [below of=10] {$\ro$};
-    \node[main node] (12) [below of=11] {$\nopermo$};
+    \node[main node] (rwlxo) [below right of=rwlx,yshift=-0.1cm] {$\rwlxo$};
+    \node[main node] (rwxo) [below of=rwlxo] {$\rwxo$};
+    \node[main node] (rxo) [below of=rwxo] {$\rxo$};
+    \node[main node] (eo) [below of=rxo] {$\entero$};
 
-    \node[main node] (8) [above left of=9] {$\rwl$};
-    \node[main node] (4) [below of=8] {$\rw$};
-    \node[main node] (5) [below of=4] {$\readonly$};
-    \node[main node] (6) [below of=5] {$\noperm$};
+    \node[main node] (rwl) [below left of=rwlx,xshift=-0.5cm,yshift=-0.3cm] {$\rwl$};
+    \node[main node] (rw) [below of=rwl] {$\rw$};
+    \node[main node] (r) [below of=rw] {$\readonly$};
+    \node[main node] (0) [below of=r] {$\noperm$};
+
+    \node[main node] (rwlo) [below right of=rwl,yshift=-0.1cm] {$\rwlo$};
+    \node[main node] (rwo) [below of=rwlo] {$\rwo$};
+    \node[main node] (ro) [below of=rwo] {$\ro$};
+    \node[main node] (0o) [below of=ro] {$\nopermo$};
 
     \path[every node/.style={font=\sffamily\small}]
-    (7) edge (1)
-    (8) edge (4)
-    (1) edge (2)
-    (2) edge (3)
-    (4) edge (5)
-    (5) edge (6)
+    (rwlxo) edge (rwxo)
+    (rwxo) edge (rxo)
+    (rxo) edge (eo)
 
-    (9) edge (10)
-    (10) edge (11)
-    (11) edge (12)
+    (rwl) edge (rw)
+    (rw) edge (r)
+    (r) edge (0)
 
-    (9) edge (7)
-    (10) edge (1)
-    (11) edge (2)
-    (12) edge (3)
+    (rwlo) edge (rwo)
+    (rwo) edge (ro)
+    (ro) edge (0o)
 
-    (9) edge (8)
-    (10) edge (4)
-    (11) edge (5)
-    (12) edge (6);
+    (rwlx) edge[dashed] (rwx)
+    (rwx) edge[dashed] (rx)
+    (rx) edge[dashed] (e)
+
+    (rwlo) edge (rwlxo)
+    (rwo) edge (rwxo)
+    (ro) edge (rxo)
+    (0o) edge (eo)
+
+    (rwlo) edge (rwl)
+    (rwo) edge (rw)
+    (ro) edge (r)
+    (0o) edge (0)
+
+    (rwl) edge (rwlx)
+    (rw) edge[dashed] (rwx)
+    (r) edge[dashed] (rx)
+    (0) edge[dashed] (e)
+
+    (rwlxo) edge (rwlx)
+    (rwxo) edge[dashed] (rwx)
+    (rxo) edge[dashed] (rx)
+    (eo) edge[dashed] (e);
   \end{tikzpicture}
@@
 \end{figure}
+
+\subsection{Operational Semantics}
+\subsubsection{Notes}
+Genrally:
+\begin{itemize}
+\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
+\end{itemize}
+
+Source language:
+\begin{itemize}
+\item No jumps to enter (call reserved for this)
+\item Variable length instructions that match the length of the compiled instructions
+  \begin{itemize}
+  \item Opaque capabilities do not hide the length of a program (especially an issue if we ever hope to have error recovery).
+  \item This is needed for correctness. See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
+  \end{itemize}
+\end{itemize}
+
+Target language:
+\begin{itemize}
+\item 
+\end{itemize}
+
+\subsubsection{Helpful functions}
+\[
+  \updPcAddr \defeq \dots
+\]
+
+\[
+  \updPcPerm \defeq \dots
+\]
+
+\[
+  \clearms \defeq \dots
+\]
+
+\[
+  \clearlocreg \defeq \dots
+\]
+
+\subsubsection{Source Language}
+
+\begin{align*}
+  \ssem{\sjmp{r}} = & 
+    \begin{cases}
+      \Phi
+      \begin{aligned}[t]
+        & \update{\sreg.\pcreg}{\opc} \\
+        & \update{\smsstk}{\ms_\stk' \uplus \clearms{\ms_\stk}}\\
+        & \update{\sastk}{a_\stk'}\\
+        & \update{\sstk}{\stk'}
+      \end{aligned} & \Phi(r) = \retptr \text{ and }
+      \Phi.\sstk = (\opc,\ms_\stk',a_\stk') :: \stk'\\
+      & \\
+      \updPc{\updPcPerm{\Phi(r)}} & \text{otherwise}
+    \end{cases} \\
+\end{align*}
+\begin{align*}
+  \ssem{\scall{r}{\rn}} = & 
+    \begin{cases}
+      (\mem,\reg',\stk',\ms_\stk') & \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and }\\
+      \\
+      \failed & \text{otherwise}
+    \end{cases}
+\end{align*}
+
+\subsubsection{Target Language}
 
@@
 \section{Examples}
-\subsection{Capability opacity}
+\subsection{Capability Opacity}
+\label{subsec:capability-opacity}
+This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions.
+The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.
+
 The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness

### 2017-03-23 fully-abstract-comp/technical-report.tex
notes

@@
 \begin{itemize}
-\item No jumps to enter (call reserved for this)
 \item Variable length instructions that match the length of the compiled instructions
@@
   \end{itemize}
+\item Programs have undefined behavior when passing ret/stk pointers.
 \end{itemize}
@@
 \section{Back translation}
-We want to do the back translation by interpreting the programs.
+The back translation is an embedding of source language into .
+
+\section{Notes}
+\subsection{Leuven stay conclusions}
+\begin{enumerate}
+\item Enter capabilities replaces by sealed code/data pairs \label{item:first-point}
+  \begin{itemize}
+  \item To allow us to forbid dynamic code generation
+  \end{itemize}
+\item Conditional full-abstraction
+  \begin{itemize}
+  \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction.
+(Blame idea: undefined as undef of current pc, so current pc can be checked for whether it was the trusted code).
+  \item Avoids dynamic checks to protect trusted code against itself.
+  \item Possible due to point \ref{item:first-point}.
+  \end{itemize}
+\item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
+  \begin{itemize}
+  \item Allows backtranslation to be embedding into source language.
+    \begin{itemize}
+    \item Interpreter for backtranslation is not able to accurately replicate code length.
+    \end{itemize}
+  \end{itemize}
+\item Worlds similar to CSF paper, but invariants on seals as well.
+  \begin{itemize}
+  \item Invariants says what can be sealed with a certain seals.
+  \end{itemize}
+\item Linker: resolves symbols
+  \begin{itemize}
+  \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
+  \item Import refs
+  \item Fresh seal requirement
+  \end{itemize}
+\item Components are memory segments and symbols.
+\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, se example in subsection~\ref{subsec:capability-opacity}).
+  \begin{itemize}
+  \item Notice x86 allows instructions of up to infinite length, so this is not weird.
+  \end{itemize}
+\item Have opaque capabilities.
+  \begin{itemize}
+  \item Without opaque capabilities, stack consumption could be inferred through stack pointer index etc.
+  \end{itemize}
+\item Considered PORs for transition systems to: (a->b => a=a and b=b)
+  \begin{itemize}
+  \item Define one region that governs the stack
+  \item Make the regions more general
+  \item For now, we have not done this: instead we stick with the transition systems we have and have to regions governing the stack reflecting the CSF work.
+  \end{itemize}
+\end{enumerate}
 \end{document}

### 2017-04-05 csf-2017/paper.tex
missing word
Added missing updatePc to load in table of operational semantics
swapped arguments to move in g1 example (review comment)

@@
   \end{align*}
-  \begin{tabular}{|c|p{4.3cm}|p{9.9cm}|}
+  \begin{tabular}{|c|p{3.5cm}|p{10.7cm}|}
     \hline
@@
     \hline
-    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)})$&\\
+    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{w})$&$w = \memreg(r_2)$\\
     \hline
-    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\updateReg{r_1}{\memheap(\addr)}$&$\memreg(r_2) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$\footnotesize{ and }$\start \leq \addr \leq \addrend$\\
+    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ \footnotesize{ and }$w = \memheap(\addr)$\footnotesize{ and }$\start \leq \addr \leq \addrend$\newline \footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
     \hline
@@
     \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline
-    \footnotesize{otherwise }$\var{newPc} = \memreg(r)$\\
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline \footnotesize{otherwise }$\var{newPc} = \memreg(r)$\\
     \hline
@@
 \begin{description}[font=\normalfont\itshape]
-\item[At program start-up] A local $\rwlx$ stack pointer in register $r_\stk$. No global write-local capabilities.
+\item[At program start-up] A local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local capabilities.
 \item[Before returning to the adversary] Clear non-return-value registers. Clear the part of the stack we had access to.
@@
      store $r_2$ 0
-     move $\pcreg$ $r_3$
+     move $r_3$ $\pcreg$
      lea $r_3$ $\var{offset}$

### 2017-05-04 popl-2018/paper.tex
popl paper first commit. csf paper copied to new format and it compiles (but it needs to be tidied up).

--- 2017-04-05 csf-2017/paper.tex
@@
-\documentclass[compsoc,conference,letterpaper,fleqn]{IEEEtran}
+%% For double-blind review submission
+\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
+%% For single-blind review submission
+%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
+%% For final camera-ready submission
+%\documentclass[acmlarge]{acmart}\settopmatter{}
+
+%% Note: Authors migrating a paper from PACMPL format to traditional
+%% SIGPLAN proceedings format should change 'acmlarge' to
+%% 'sigplan,10pt'.
+
+%% Some recommended packages.
+\usepackage{booktabs}   %% For formal tables:
+                        %% http://ctan.org/pkg/booktabs
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
 
@@
 \usepackage{stmaryrd}
-\usepackage[numbers]{natbib}
+% \usepackage[numbers]{natbib}
 \usepackage{url}
@@
 
-\ifCLASSOPTIONcompsoc
-\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\else
-\usepackage[caption=false,font=footnotesize]{subfig}
-\fi
-
-% Include the following to reduce room around display equations
-%\setlength{\abovedisplayskip}{3pt}
-%\setlength{\belowdisplayskip}{3pt}
-%\setlength{\abovedisplayshortskip}{3pt}
-%\setlength{\belowdisplayshortskip}{3pt}
-
-% %make latex preview-mode work with natbib...
-% \usepackage[displaymath,floats,graphics,textmath,footnotes]{preview}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
+% \ifCLASSOPTIONcompsoc
+% \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+% \else
+% \usepackage[caption=false,font=footnotesize]{subfig}
+% \fi
+
+% \newtheorem{theorem}{Theorem}
+% \newtheorem{lemma}{Lemma}
 
@@
 \newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
-
 % Math notation
@@
 \newcommand{\id}{\var{id}}
-\newcommand{\undefined}{\mathit{undefined}}
+% \newcommand{\undefined}{\mathit{undefined}}
 
@@
 \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
-
 % domi: IEEEtran.cls sets font size to 10bp, apparently.
@@
 
-\newlength{\oldmathindent}
-\newenvironment{withmathindent}[1]{\setlength{\oldmathindent}{\mathindent}\setlength{\mathindent}{#1}}{\setlength{\mathindent}{\oldmathindent}}
+% \newlength{\oldmathindent}
+% \newenvironment{withmathindent}[1]{\setlength{\oldmathindent}{\mathindent}\setlength{\mathindent}{#1}}{\setlength{\mathindent}{\oldmathindent}}
+
+\makeatletter\if@ACM@journal\makeatother
+%% Journal information (used by PACMPL format)
+%% Supplied to authors by publisher for camera-ready submission
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{1}
+\acmArticle{1}
+\acmYear{2017}
+\acmMonth{1}
+\acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+\else\makeatother
+%% Conference information (used by SIGPLAN proceedings format)
+%% Supplied to authors by publisher for camera-ready submission
+\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
+\acmYear{2017}
+\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+\acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+\fi
+
+%% Copyright information
+%% Supplied to authors (based on authors' rights management selection;
+%% see authors.acm.org) by publisher for camera-ready submission
+\setcopyright{none}             %% For review submission
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\copyrightyear{2017}           %% If different from \acmYear
+
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+%% Citation style
+%% Note: author/year citations are required for papers published as an
+%% issue of PACMPL.
+\citestyle{acmauthoryear}   %% For author/year citations
 
 \begin{document}
-\setlength{\mathindent}{.2cm}
-
-\title{Reasoning about a Capability Machine with Local Capabilities\\
- Provably Safe Stack and Return Pointer Management (without OS Support)}
-
-\author{%
-  % \IEEEauthorblockN{Lau~Skorstengaard}
-  % \IEEEauthorblockA{Aarhus University, Denmark\\
-  %   Email: lau@cs.au.dk} \and
-  % \IEEEauthorblockN{Dominique~Devriese}
-  % \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
-  %   Email: dominique.devriese@cs.kuleuven.be} \and
-  % \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
-  %   Email: birkedal@cs.au.dk}
+
+%% Title information
+\title[Short Title]{Full Title}         %% [Short Title] is optional;
+                                        %% when present, will be used in
+                                        %% header instead of Full Title.
+\titlenote{with title note}             %% \titlenote is optional;
+                                        %% can be repeated if necessary;
+                                        %% contents suppressed with 'anonymous'
+\subtitle{Subtitle}                     %% \subtitle is optional
+\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+                                        %% can be repeated if necessary;
+                                        %% contents suppressed with 'anonymous'
+
+%% Author information
+%% Contents and number of authors suppressed with 'anonymous'.
+%% Each author should be introduced by \author, followed by
+%% \authornote (optional), \orcid (optional), \affiliation, and
+%% \email.
+%% An author may have multiple affiliations and/or emails; repeat the
+%% appropriate command.
+%% Many elements are not rendered, but should be provided for metadata
+%% extraction tools.
+
+%% Author with single affiliation.
+\author{Lau~Skorstengaard}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Aarhus University}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{Aarhus}
+  \state{State1}
+  \postcode{8210}
+  \country{Denmark}
 }
-\maketitle
-
+\email{lau@cs.au.dk}          %% \email is recommended
+
+\author{Dominique~Devriese}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{iMinds-DistriNet}              %% \department is recommended
+  \institution{KU Leuven}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{-}
+  \state{-}
+  \postcode{-}
+  \country{Belgium}
+}
+\email{dominique.devriese@cs.kuleuven.be}          %% \email is recommended
+
+\author{Lars~Birkedal}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Aarhus University}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{Aarhus}
+  \state{State1}
+  \postcode{8210}
+  \country{Denmark}
+}
+\email{birkedal@cs.au.dk}          %% \email is recommended
+
+%% Paper note
+%% The \thanks command may be used to create a "paper note" ---
+%% similar to a title note or an author note, but not explicitly
+%% associated with a particular element.  It will appear immediately
+%% above the permission/copyright statement.
+\thanks{with paper note}                %% \thanks is optional
+                                        %% can be repeated if necesary
+                                        %% contents suppressed with 'anonymous'
+
+%% Abstract
+%% Note: \begin{abstract}...\end{abstract} environment must come
+%% before \maketitle command
 \begin{abstract}
@@
 
-% \section*{Terminology ?}
-
-% \emph{capability safe}: 
-% Our unary logical relation is a semantic \emph{definition}
-% of which values for the
-% pc / register-file / word are capability safe.
-
-% \emph{adversary}: It doesn't sound ``sexy'' to constantly say ``an untrusted
-% piece of code'' should we adopt the convention that adversary refers
-% to an unstrusted piece of code?
-
-% \emph{memory}  we don't make a distinction between memory and memory
-% segments in the text  (to do so properly, we would need to explain why
-% assumptions
-% are reasonable, given that a real machine has finite memory, but we
-% need infinite domain for malloc spec, etc.)
+%% 2012 ACM Computing Classification System (CSS) concepts
+%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011008</concept_id>
+<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003456.10003457.10003521.10003525</concept_id>
+<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~General programming languages}
+\ccsdesc[300]{Social and professional topics~History of programming languages}
+%% End of generated code
+
+%% Keywords
+%% comma separated list
+% \keywords{keyword1, keyword2, keyword3}  %% \keywords is optional
+
+%% \maketitle
+%% Note: \maketitle command must come after title commands, author
+%% commands, abstract environment, Computing Classification System
+%% environment and commands, and keywords command.
+\maketitle
 
@@
   \centering
-  \begin{withmathindent}{0cm}
+%  \begin{withmathindent}{0cm}
     \begin{align*}
@@
     \end{align*}
-  \end{withmathindent}
+%  \end{withmathindent}
   \caption{Logical Relation}
@@
 
-%\section{Conclusion}
-%\label{sec:conclusion}
-
-\section*{Acknowledgements}
-\label{sec:acknowledgements}
-
-This research was supported in part by the ModuRes Sapere Aude Advanced Grant from The Danish Council for Independent Research for the Natural Sciences (FNU).
-Dominique Devriese holds a postdoctoral fellowship from the Research Foundation - Flanders (FWO).
-
-\bibliographystyle{IEEEtranN}
-% \bibliographystyle{plainnat}
-%\bibliographystyle{plain}
-\bibliography{IEEEabrv,references}
+%% Acknowledgments
+\begin{acks}                            %% acks environment is optional
+                                        %% contents suppressed with 'anonymous'
+  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+  %% acknowledge financial support and will be used by metadata
+  %% extraction tools.
+  This material is based upon work supported by the
+  \grantsponsor{GS100000001}{National Science
+    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+  conclusions or recommendations expressed in this material are those
+  of the author and do not necessarily reflect the views of the
+  National Science Foundation.
+\end{acks}
+
+%% Bibliography
+\bibliography{references}
+
+%% Appendix
+% \appendix
+% \section{Appendix}
+
+% Text of appendix \ldots
 
 \end{document}
-\grid

### 2017-05-24 csf-2017/paper.tex
n

--- 2017-03-15 csf-2017/paper.tex
@@
 \author{%
-  % \IEEEauthorblockN{Lau~Skorstengaard}
-  % \IEEEauthorblockA{Aarhus University, Denmark\\
-  %   Email: lau@cs.au.dk} \and
-  % \IEEEauthorblockN{Dominique~Devriese}
-  % \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
-  %   Email: dominique.devriese@cs.kuleuven.be} \and
-  % \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
-  %   Email: birkedal@cs.au.dk}
+  \IEEEauthorblockN{Lau~Skorstengaard}
+  \IEEEauthorblockA{Aarhus University, Denmark\\
+    Email: lau@cs.au.dk} \and
+  \IEEEauthorblockN{Dominique~Devriese}
+  \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
+    Email: dominique.devriese@cs.kuleuven.be} \and
+  \IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
+    Email: birkedal@cs.au.dk}
 }

### 2017-05-24 fully-abstract-comp/technical-report.tex

@@
 \newcommand{\src}[1]{{\sourcecolor{} #1}}
-\newcommand{\targetcolortext}{green}
-\newcommand{\targetcolor}[1]{\color{green}}
+\newcommand{\targetcolortext}{black}
+\newcommand{\targetcolor}[1]{\color{black}}
 \newcommand{\trg}[1]{{\targetcolor{} #1}}
@@
 \newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
+\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
+\newcommand{\tcseal}[2]{\twoinstr{\trg{cseal}}{#1}{#2}}
 
@@
 \newcommand{\update}[2]{[#1 \mapsto #2]}
-\newcommand{\updReg}[3][\Phi]{#1\update{\sreg.#2}{#3}}
+\newcommand{\updReg}[3][\Phi]{#1\update{\reg.#2}{#3}}
 \newcommand{\updPc}[2][\Phi]{\updReg[#1]{\pcreg}{#2}}
@@
 %%% Source dom
-\newcommand{\sourcedom}[1]{\mathrm{#1}}
-\newcommand{\sRegName}{\sourcedom{RegisterName}}
-\newcommand{\sAddr}{\sourcedom{Addr}}
-\newcommand{\sPerm}{\sourcedom{Perm}}
-\newcommand{\sCap}{\sourcedom{Cap}}
-\newcommand{\sWord}{\sourcedom{Word}}
-\newcommand{\sMem}{\sourcedom{Memory}}
-\newcommand{\sReg}{\sourcedom{RegisterFile}}
-\newcommand{\sStk}{\sourcedom{Stack}}
-\newcommand{\sConf}{\sourcedom{Conf}}
-\newcommand{\sExecConf}{\sourcedom{ExecConf}}
-\newcommand{\sGlobal}{\sourcedom{Global}}
-\newcommand{\sMemSeg}{\sourcedom{MemorySegment}}
-\newcommand{\sStkFrame}{\sourcedom{StackFrame}}
-\newcommand{\sStack}{\sourcedom{Stack}}
+\newcommand{\shareddom}[1]{\mathrm{#1}}
+\newcommand{\RegName}{\shareddom{RegisterName}}
+\newcommand{\Addr}{\shareddom{Addr}}
+\newcommand{\Seal}{\shareddom{Seal}}
+\newcommand{\Perm}{\shareddom{Perm}}
+\newcommand{\Caps}{\shareddom{Cap}}
+\newcommand{\SealableCaps}{\shareddom{SealableCap}}
+\newcommand{\Word}{\shareddom{Word}}
+\newcommand{\Mem}{\shareddom{Memory}}
+\newcommand{\Reg}{\shareddom{RegisterFile}}
+\newcommand{\Stk}{\shareddom{Stack}}
+\newcommand{\Conf}{\shareddom{Conf}}
+\newcommand{\ExecConf}{\shareddom{ExecConf}}
+\newcommand{\Global}{\shareddom{Global}}
+\newcommand{\MemSeg}{\shareddom{MemorySegment}}
+\newcommand{\StkFrame}{\shareddom{StackFrame}}
+\newcommand{\Stack}{\shareddom{Stack}}
 
-\newcommand{\smem}{\sourcedom{mem}}
-\newcommand{\sreg}{\sourcedom{reg}}
-\newcommand{\sstk}{\sourcedom{stk}}
-\newcommand{\smsstk}{\sourcedom{ms_{stk}}}
-\newcommand{\sstkframe}{\sourcedom{frame}}
-\newcommand{\sopc}{\sourcedom{opc}}
-\newcommand{\sastk}{\sourcedom{a_{stk}}}
-\newcommand{\sperm}{\sourcedom{perm}}
-\newcommand{\sgl}{\sourcedom{g}}
-\newcommand{\sbase}{\sourcedom{base}}
-\newcommand{\send}{\sourcedom{end}}
-\newcommand{\saddr}{\sourcedom{a}}
-\newcommand{\scap}{\sourcedom{c}}
-\newcommand{\sms}{\sourcedom{ms}}
+\newcommand{\scbnf}{\shareddom{sc}}
+\newcommand{\cbnf}{\shareddom{c}}
+\newcommand{\permbnf}{\shareddom{perm}}
+\newcommand{\addrbnf}{\shareddom{a}}
+\newcommand{\basebnf}{\shareddom{base}}
+\newcommand{\aendbnf}{\shareddom{end}}
+\newcommand{\glbnf}{\shareddom{gl}}
+\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
+\newcommand{\sealendbnf}{\sigma_\shareddom{end}}
 
-\newcommand{\stkptr}[4]{\sourcedom{stack\text{-}ptr}(#1,#2,#3,#4)}
-\newcommand{\retptr}{\sourcedom{ret\text{-}ptr}}
+\newcommand{\sstk}{\shareddom{stk}}
+\newcommand{\smsstk}{\shareddom{ms_{stk}}}
+\newcommand{\sstkframe}{\shareddom{frame}}
+\newcommand{\sopc}{\shareddom{opc}}
+\newcommand{\sastk}{\shareddom{a_{stk}}}
+\newcommand{\perm}{\shareddom{perm}}
+\newcommand{\gl}{\shareddom{g}}
+\newcommand{\base}{\shareddom{base}}
+\newcommand{\aend}{\shareddom{end}}
+\newcommand{\addr}{\shareddom{a}}
+\newcommand{\scap}{\shareddom{c}}
+\newcommand{\sms}{\shareddom{ms}}
+
+\newcommand{\stkptr}[1]{\mathrm{stack\text{-}ptr}(#1)}
+\newcommand{\retptr}{\mathrm{ret\text{-}ptr}}
+\newcommand{\retptrd}{\mathrm{ret\text{-}ptr\text{-}data}}
+\newcommand{\retptrc}{\mathrm{ret\text{-}ptr\text{-}code}}
+
+\newcommand{\seal}[1]{\shareddom{seal}(#1)}
+\newcommand{\sealed}[1]{\shareddom{sealed}(#1)}
 
@@
 \newcommand{\rO}{\mathrm{r}_\mathrm{ret}}
+\newcommand{\rret}{\rO}
+\newcommand{\rdata}{\mathrm{r}_\mathrm{data}}
 
@@
 \begin{document}
-\section{TODO}
-\begin{itemize}
-\item Linking
-\item Component
-\item Cross-language LR
-\end{itemize}
-
 \section{The two capability machines}
@@
   \begin{array}{rrcl}
-   \saddr,\sbase \in & \sAddr & \defeq & \nats \\
-    &\sWord & \defeq & \ints \uplus \sCap\\
-   \sperm \in& \sPerm & \defbnf & \dots \\
-    \sgl \in&\sGlobal & \defbnf & \glob \mid \local \\
-    &\sCap &::= & \left\{((\sperm,\sgl),\sbase,\send,\saddr) \mid \send \in \sAddr \cup \{\infty\}\right\} {\sourcecolor{} \uplus}\\
-    & & & {\sourcecolor{} \{ \stkptr{\sperm}{\sbase}{\send}{\saddr} \}\uplus}\\ 
-    & & & {\sourcecolor{} \{ \retptr \}}\\ 
-    &\sRegName & \defbnf &\pcreg \mid \rO \mid \rstk \mid \dots \\
-    &\sReg & \defeq & \sRegName \fun \sWord\\
-    & \sMem & \defeq & \sAddr \fun \sWord \\
-    & \sMemSeg & \defeq & \sAddr \parfun \sWord \\
-    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \sStkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \sCap \times \sMemSeg \times \sAddr}\\
-    {\sourcecolor{} \sstk \in}& {\sourcecolor{} \sStack} & {\sourcecolor{} \defeq} & {\sourcecolor{} \sStkFrame^*} \\
-    \Phi \in & \sExecConf & \defeq & \sMem \times \sReg {\sourcecolor{} \; \times \; \sStack \times \sMemSeg }\\
-    &\sConf & \defeq & \sExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \sMem)
+   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
+    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
+    &\Word & \defeq & \ints \uplus \Caps\\
+    \permbnf \in& \Perm & \defbnf & \dots \\
+    &\glbnf & \defbnf & \glob \mid \local \\
+    & \aendbnf & \in & \Addr \uplus \{\infty \} \\
+    &  \sealendbnf & \in & \Seal \uplus \{\infty \} \\
+    &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
+    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\     &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
+    &\RegName & \defbnf &\pcreg \mid \rO \mid \rstk \mid \rdata \mid \dots \\
+    &\Reg & \defeq & \RegName \fun \Word\\
+    &\Mem & \defeq & \Addr \fun \Word \\
+    &\MemSeg & \defeq & \Addr \parfun \Word \\
+    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
+    {\sourcecolor{} \sstk \in}& {\sourcecolor{} \Stack} & {\sourcecolor{} \defeq} & {\sourcecolor{} \StkFrame^*} \\
+    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
+    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
   \end{array}
 \]
-The target language domains are all the non blue parts in the above.
+The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
+\begin{itemize}
+\item $\glbnf$ defines domain $\Global$
+\item $\scbnf$ defines domain $\SealableCaps$
+\item $\cbnf$ defines domain $\Caps$
+\end{itemize}
 
@@
   \begin{array}{rcl}
-    \Phi(r) & \defeq & \Phi.\sreg(r)
+    \Phi(r) & \defeq & \Phi.\reg(r)
   \end{array}
 \]
-where $r\in \sRegName$
+where $r\in \RegName$
 
 \subsection{Syntax}
-The target machine is[dashed] a simple capability machine with memory capabilities, local capabilities and enter capabilities.
+The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle} (Inspired by CHERI).
 The syntax of the instructions of the target machine is defined as follows:
@@
       & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
-      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}
+      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}
 \end{array}
@@
 
-The source machine is also a capability machine with memory capabilities, local capabilities and enter capabilities.
-Unlike the target machine, the source machine is going to have a built in stack.
+The source machine is also a capability machine with memory capabilities, local capabilities and sealed capabilities.
+Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities.
 The syntax of the source machine language is as follows:
@@
     n & \in & \nats \\
-    \src{r} &\in &  \sRegName \setminus \{\rO,\rstk\}\\
+    \src{r} &\in &  \RegName \setminus \{\rO,\rstk\}\\
     \src{\rn} &::= & \src{r} \mid n \\
-    \src{i} & ::= &  \trg{i} \mid \scall{\src{r}}{\src{\rn}}
+    \src{i} & ::= &  \trg{i} \mid \scall{\src{r}}{\src{r}}
   \end{array}
 \]
-There is one suntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
+There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
+
 \subsection{Permissions}
+\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.}
 \begin{figure}[!h]
@@
 \subsubsection{Source Language}
-
+\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update.}
 \begin{align*}
@@
       \begin{aligned}[t]
-        & \update{\sreg.\pcreg}{\opc} \\
+        & \update{\reg.\pcreg}{\opc} \\
         & \update{\smsstk}{\ms_\stk' \uplus \clearms{\ms_\stk}}\\
@@
 \begin{align*}
-  \ssem{\scall{r}{\rn}} = & 
+  \ssem{\scall{r_1}{r_2}} = & 
     \begin{cases}
@@
 \end{align*}
+\begin{itemize}
+\item $\scall{}{}$ does not do argument spilling anymore. We leave it to the caller to push the arguments for spilling to the stack and update the stack pointer accordingly.
+\end{itemize}
 
@@
 \end{lstlisting}
-distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
+\footnote{At this time $\scall{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
 We would not have enough instructions to set up a proper return pointer for the compiled return to use.

### 2017-05-24 local-capability-machine/local-capability-machine.tex

@@
      store $r_2$ 0
-     move $\pcreg$ $r_3$
+     move $r_3$ $\pcreg$
      lea $r_3$ $\dots$

### 2017-05-24 popl-2018/paper.tex
Added title and started changing syntax figure.
Updated syntax figure to still be very compact and fit in new format (readability questionable).
Figure with excerpt op sem updated to fit in column.

@@
 %% Title information
-\title[Short Title]{Full Title}         %% [Short Title] is optional;
+\title{Reasoning about a Capability Machine with Local Capabilities}         %% [Short Title] is optional;
                                         %% when present, will be used in
                                         %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
+%\titlenote{with title note}             %% \titlenote is optional;
                                         %% can be repeated if necessary;
                                         %% contents suppressed with 'anonymous'
-\subtitle{Subtitle}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+\subtitle{Provably Safe Stack and Return Pointer Management (without OS Support)}                     %% \subtitle is optional
+%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                         %% can be repeated if necessary;
@@
 %% above the permission/copyright statement.
-\thanks{with paper note}                %% \thanks is optional
+%\thanks{with paper note}                %% \thanks is optional
                                         %% can be repeated if necesary
@@
 \begin{figure}
-  \begin{align*}
-    \addr   &\in& \Addrs &\isdef \nats\\
-    w &\in&\Words &\isdef \ints + \Caps \\
-    \perm   &\in& \Perms &::= \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid \exec\mid \entry\mid \rwx\mid \rwlx\\
-    \gl&\in&\Globals & ::= \glob \mid \local \\
-     && \Caps  &::= \left\{
-                       \begin{multlined}
-                         ((\perm,\gl),\start,\addrend,\addr)\mid\\
+\[
+  \begin{array}{rlrcl p{2cm}rlrcl}
+    \addr   &\in& \Addrs &\isdef& \nats & & r &\in& \RegName&\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
+    w &\in&\Words &\isdef& \ints + \Caps & & \reg &\in& \Regs  &\isdef& \RegName \rightarrow \Words\\
+    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m&\in& \Heaps &\isdef& \Addrs \rightarrow \Words \\
+    & & & &\exec\mid \entry\mid \rwx\mid \rwlx & & \Phi &\in& \ExecConfs  &\isdef& \Regs \times \Heaps \\
+    \gl&\in&\Globals & ::= &\glob \mid \local & & \ms &\in& \MemSegments &::= & \Addrs \parfun \Words \\
+    &&\Confs &::=& \multicolumn{5}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
+     && \Caps  &::= & \multicolumn{5}{l}{\left\{
+                         ((\perm,\gl),\start,\addrend,\addr)\mid
                          \start,\addr\in\Addrs,\addrend \in
-                         \Addrs\cup \{\infty\}
-                       \end{multlined} \right\}\\
-    r       &\in& \RegName&\bnfdef \pcreg\mid r_0\mid r_1\mid\ldots\\
-    \reg &\in& \Regs  &\isdef \RegName \rightarrow \Words\\
-    m&\in& \Heaps &\isdef \Addrs \rightarrow \Words \\
-    \Phi    &\in& \ExecConfs  &\isdef \Regs \times \Heaps \\
-    &&\Confs &::= \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps \\
-    \ms     &\in& \MemSegments &::= \Addrs \parfun \Words
-  \end{align*}
+                         \Addrs\cup \{\infty\}\right\}}
+  \end{array}
+\]
   \begin{equation*}
@@
                  \load{\lv}{\hv} \mid 
-                 \store{\hv}{\rv} \mid  \\
-           &   & \plus{\lv}{\rv}{\rv} \mid 
+                 \store{\hv}{\rv} \mid \plus{\lv}{\rv}{\rv} \mid 
                  \minus{\lv}{\rv}{\rv} \mid 
@@
                  \getp{\lv}{\lv} \mid 
-                 \getl{\lv}{\lv} \mid\\ 
-           &   & \getb{\lv}{\lv} \mid
+                 \getl{\lv}{\lv} \mid \getb{\lv}{\lv} \mid
                  \gete{\lv}{\lv} \mid
@@
     \begin{cases}
-      \sem{\decode(n)}(\Phi) & \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }}\\
-      \failed                                 & \text{\footnotesize{otherwise}}
+      \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
+      \begin{array}[t]{l}
+\text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$}\\ \text{\;\;and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }
+      \end{array}
+\\
+      \failed                                 & \text{otherwise}
     \end{cases}
-  \end{equation*}\vspace{-.3cm}%
+  \end{equation*}
   \begin{align*}
@@
                            \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
-                             \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$ and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}\\
-                             \failed & \text{\footnotesize{otherwise}}
+                             \text{if $\memreg(\pcreg) = \stdcap$ and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}\\
+                             \failed & \text{otherwise}
                          \end{cases}
   \end{align*}
-  \begin{tabular}{|c|p{3.5cm}|p{10.7cm}|}
+  \begin{tabular}{|c|p{3.9cm}|p{8.3cm}|}
     \hline
@@
     \hline
-    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ \footnotesize{ and }$w = \memheap(\addr)$\footnotesize{ and }$\start \leq \addr \leq \addrend$\newline \footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
+    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and $w = \memheap(\addr)$ and$\start \leq \addr \leq \addrend$\newline and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
     \hline
-    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$\footnotesize{ and }$(\perm',g') = \decodePermPair{\memreg(r_2)}$
-                                                      \footnotesize{ and }$(\perm',g') \sqsubseteq (\perm,g)$\footnotesize{ and } $w =((\perm',g'),\start,\addrend,\addr)$\\
+    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$ and $(\perm',g') = \decodePermPair{\memreg(r_2)}$
+                                                      and $(\perm',g') \sqsubseteq (\perm,g)$ and  $w =((\perm',g'),\start,\addrend,\addr)$\\
     \hline
@@
     \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& \footnotesize{if }$\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$\footnotesize{, then} $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline \footnotesize{otherwise }$\var{newPc} = \memreg(r)$\\
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline otherwise $\var{newPc} = \memreg(r)$\\
     \hline
-    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$\footnotesize{ and }$\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$ \footnotesize{ and } $\start \leq \addr \leq \addrend$\footnotesize{ and }$\var{w} = \memreg(r_2)$
-                                                                                       \footnotesize{and if }$\var{w} = ((\_,\local),\_,\_,\_)$\footnotesize{, then  }$\perm \in \{\rwlx,\readwritel \}$\\
+    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
+                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
     \hline
@@
     \hline
-    \_&$\failed$&\footnotesize{otherwise}\\
+    \_&$\failed$&otherwise\\
     \hline
@@
         \begin{array}[t]{l}
-          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\
-          \;\;\text{\footnotesize{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}}\\
-          \;\;\text{\footnotesize{and $\perm \in \{ \exec,\rwx, \rwlx \}$} }
+          \text{if $\memreg(\pcreg) = \stdcap$}\\
+          \;\;\text{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}\\
+          \;\;\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$} 
         \end{array}\\
-        \failed                & \text{\footnotesize{otherwise}}
+        \failed                & \text{otherwise}
       \end{cases}
@@
         ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
-        w & \text{\footnotesize{otherwise}}
+        w & \text{otherwise}
       \end{cases}\\
@@
         \begin{array}[t]{l}
-          \text{\footnotesize{if $\memreg(\pcreg) = \stdcap$}}\\ 
-          \;\;\text{\footnotesize{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}}
+          \text{if $\memreg(\pcreg) = \stdcap$}\\ 
+          \;\;\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
         \end{array}\\
-        \failed & \text{\footnotesize{otherwise}}
+        \failed & \text{otherwise}
       \end{cases}
@@
   \begin{align*}
-      \xi &\; : \; \Wor \cong \blater (\nats \finparfun \Regions)\\
-      \Regions & = \{\revoked\} \uplus \\
-                 &\;\quad \begin{multlined}
-                     \{\temp\} \times \States \times \Rels \times \\
-                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
-                   \end{multlined} \uplus \\
-                 &\;\quad \begin{multlined}
-                     \{\perma\} \times \States \times \Rels \times \\
-                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
-                   \end{multlined}
+      \xi \; : \; & \Wor \cong \blater (\nats \finparfun \Regions)\\
+      \Regions  = & \; \{\revoked\} \uplus \\
+               & \; \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))\uplus \\
+               & \; \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
     \end{align*}
@@
   &\memSat{\ms}{W}
-    \text{ iff }\\
-  &\quad \left\{\begin{aligned}
-        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{\footnotesize{ and }} \\[-.2cm]
-        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp\\
-        &\qquad W(r) = (\_,s,\_,\_,H) \text{\footnotesize{ and }} \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
+    \text{ iff }\quad \left\{\begin{aligned}
+        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.2cm]
+        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp W(r) = (\_,s,\_,\_,H) \text{ and } \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
       \end{aligned}\right.
@@
             & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
-            & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{\footnotesize{ and }} \\
+            & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \\
             & \;\; \heapSat[\hs']{n-i}{W'}
@@
             &\npair{\stdcap[(\readwrite,\gl)] } \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
              &\quad\npair{(\start,\addrend)} \in
@@
             &\npair{\stdcap[(\readwritel,\gl)] } \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
              &\quad\npair{(\start,\addrend)} \in
@@
             &\npair{\stdcap[(\exec,\gl)]} \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
              &\quad\npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
@@
             &\npair{\stdcap[(\rwx,\gl)]} \mid \\
-            &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-            &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
+            &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+            &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
             &\quad\npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
@@
             &\npair{\stdcap[(\rwlx,\gl)]} \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
-             &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
+             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+             &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
              &\quad\npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in
@@
       & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
+      & \;\; W(r) \text{ is address-stratified and} \\
       & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
@@
         \begin{aligned}[t]
-        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+        &\perm = \exec \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
       \end{aligned} \\
@@
       \begin{aligned}[t]
-        &\perm = \rwx \text{\footnotesize{ and }} \\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
+        &\perm = \rwx \text{ and } \\
+        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
         &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
@@
       \begin{aligned}[t]
-        &\perm = \rwlx \text{\footnotesize{ and }}\\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
+        &\perm = \rwlx \text{ and }\\
+        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
         &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),

### 2017-05-30 popl-2018/paper.tex
Some small updates while re-reading the paper.

@@
     \hline
-    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and $w = \memheap(\addr)$ and$\start \leq \addr \leq \addrend$\newline and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
+    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and $w = \memheap(\addr)$ \newline and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
     \hline
@@
 the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
-we have used the $\subseg{}{}{}$ to shrink its authority to the part that we are
-not using ourselves.
+we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
+that we are not using ourselves.
 
@@
 By making the return pointer local, we prevent such attacks: the adversary can
-only store local capabilities through write-local capabilities, which (because
-of our assumptions above) means: on the stack. Since the stack pointer itself is
+only store local capabilities through write-local capabilities, which means (because
+of our assumptions above): on the stack. Since the stack pointer itself is
 also local, it can also only be stored on the stack. Because we clear the part
@@
 restricted authority over the stack than the code making the call, we cannot
-hope to reconstruct the return pointer from the stack pointer owned by the
-adversary. Instead, the old stack pointer should be stored as part of the
-activation record.
+hope to reconstruct the original stack pointer from the stack pointer owned by the
+adversary. Instead, it should be stored as part of the activation record.
 
@@
 easy to provide thanks to the capability machine's fine-grained memory
-protection.
+protection\todo{explain this point better}.
 
@@
 perhaps the hardest scenario: how to provide a callback to the adversary. The
-basic idea is that we allocate blocks of memory using malloc that we fill with
-the capabilities and data that the callback needs access to, as well as some
+basic idea is that we allocate a block of memory using malloc that we fill with
+the capabilities and data that the callback needs, as well as some
 prelude instructions that load the data into registers and jumps to the right
@@
 the allocated block and pointing to the contained prelude instructions. However,
-the question that remains in this setup is: where do we get a stack pointer from when
+the question that remains in this setup is: from where do we get a stack pointer when
 the callback is invoked?
@@
 \item[Before invoking an adversary callback] Make sure callback is global.
-\item[When called by an adversary] Make sure received stack pointer has permission $\rwlx$.
+\item[When invoked by an adversary] Make sure received stack pointer has permission $\rwlx$.
 \end{description}
@@
 macros capturing the calling convention and other conveniences. All
-macros that uses the stack assumes a stack pointer in register
+macros that use the stack assume a stack pointer in register
 $r_\stk$. The macro \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
@@
 above. That is, it pushes local state (the contents of registers
-$\overline{r_{\var{priv}}}$), the necessary on stack return code,
-return pointer, stack pointer to the stack, and creates an $\entry$
+$\overline{r_{\var{priv}}}$) and the activation record (return code,
+return pointer, stack pointer) to the stack, creates an $\entry$
 return pointer, restricts the stack pointer, clears the unused part of
@@
 notice that the adversary capability is an enter capability, which the
-Fundamental Theorem says nothing about. Luckily the enter capability becomes
+Fundamental Theorem says nothing about.\todo{why doesn't it?}  Luckily the enter capability becomes
 $\exec$ after the jump and then the Fundamental Theorem applies.

### 2017-05-31 popl-2018/paper.tex
Use the right todo note command...

@@
 easy to provide thanks to the capability machine's fine-grained memory
-protection\todo{explain this point better}.
+protection\dominique{explain this point better}.
 
@@
 notice that the adversary capability is an enter capability, which the
-Fundamental Theorem says nothing about.\todo{why doesn't it?}  Luckily the enter capability becomes
+Fundamental Theorem says nothing about.\dominique{why doesn't it?}  Luckily the enter capability becomes
 $\exec$ after the jump and then the Fundamental Theorem applies.

### 2017-05-31 popl-2018/paper.tex
An attempt at an introductory section about the logical relation.

@@
 \label{sec:logical-relation}
-In order to reason about programs on the capability machine and prove properties
-such as well-bracketedness and encapsulation, we define a unary logical
-relation.
-% Intuition
-It is a semantic model of all the words that are definitely safe to give to an
-adversary. Whether a capability is safe depends heavily on the contents of the
-memory since instructions as well as capabilities may reside there. Hence we
-index the logical relation over worlds, modelling the memory (this makes our
-logical relation into a Kripke logical relation). A world consists of a finite
-set of regions, and each region defines a protocol that governs the possible
-evolution of a part of memory. A future world relation defines how worlds may
-evolve during execution (e.g. new protocols established over memory received
-from malloc, or steps in a protocol already established for a part of memory).
-In fact, as we will explain below, there are two future world relations to
-accommodate reasoning about local capabilities.
+Now that we have defined our calling convention, how can we be sure that it
+works? More concretely, suppose that we have a program that uses the convention
+in its interaction with untrusted adversary code. Can we formally prove the
+program's correctness if it relies on well-bracketed control flow and private
+state encapsulation for the interaction with the adversary? Clearly, such a
+proof should depend on a formal expression of the guarantees provided by the
+capability machine, including the specific guarantees for local capbalities.
+
+In this section, we construct such a formalisation. We make use of some
+well-studied and powerful (but non-trivial) machinery from the literature.
+Specifically, we employ a unary step-indexed Kripke logical relation with
+recursive worlds, and some additional special characteristics of our own.
+Step-indexing, Kripke logical relations and recursive worlds are techniques that
+may be familiar from lambda calculus settings, but it may not be clear to the
+reader how they apply in this more low-level assembly language. Therefore, we do
+not immediately dive into the details, but we first try to provide some informal
+intuition about how all of this machinery comes into play in our setting, in the
+next section.
+
+\subsection{Formalizing the guarantees of the capability machine}
+What differentiates a capability machine from a more standard assembly language
+is that we can bound the authority of an executing block of code, based solely
+on the capabilities it has access to. Specifically, it does not matter which
+instructions are actually executed, i.e. the bound also applies to untrusted
+adversary code that has not been inspected or modified in any way.
+
+\paragraph{Worlds}
+But what does a ``bound on the authority'' of an executing block of code mean?
+In our setting, there are no externally observable side-effects and the only
+primitive authority that code may hold is authority over memory. As such, the
+authority bounds we consider are related to memory, but in a form that is more
+fine-grained than standard read/write authority: code's authority can be bounded
+by arbitrary memory invariants that it is required to respect. Specifically, we
+will define worlds $W \in \Worlds$ that define a set of memory invariants and
+our results will express authority bounds on code as safety with respect to such
+a world, i.e. the fact that the code respects the invariants registered in
+the world.
+
+\paragraph{Safe values}
+So, let's say that we have a world $W$ expressing that the memory must contain
+value $42$ at address $0$ must remain $42$, may contain arbitrary values at
+address $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$ and
+an integer at address $100$ that may only increase over time\footnote{Indeed, we
+  will allow a notion of \emph{evolvable} invariants or protocols that can
+  express such a temporal property.}. Our main theorem will state that if the
+current register file only contains safe words (numbers or capabilities which
+preserve the invariants in $W$ under any interaction), then the execution will
+necessarily also preserve the invariants (irrespective of the instructions being
+executed).
+
+To make this more precise, we need to define the set $\stdvr(W) \in
+\powerset{\Words}$ of words that are safe w.r.t. $W$. Essentially, the set
+should only include words that preserve $W$'s invariants under any interaction,
+but should otherwise be as liberal as possible. Numbers are clearly always safe,
+as they cannot be used to break invariants. Whether a capability is safe depends
+on the authority that it grants. In the above-described world, a read capability
+for address $0$ is safe, as it can only be used to read the value $42$, which is
+itself safe. However, a write capability for address $0$ is not safe: it can be
+used to overwrite the memory at that address with a value other than $42$,
+breaking the invariant for that address.
+
+\paragraph{Step-indexing}
+More generally, we want to define that a read capability for memory range
+$[b,e]$ is safe if the world guarantees that the words at those addresses are
+themselves safe. However, this definition is cyclic: what if the world
+guarantees that the memory at address $a$ will contain a read capability for
+address $a$? The definition then just says that a read capability for address
+$a$ is safe iff the same read capability for address $a$ is safe. This form of
+cyclic reasoning is related to similar problems in languages with recursive
+types or higher-order ML-style references, and a standard solution is
+step-indexing. Essentially, the cycle is broken by defining safety up to a
+certain number of interaction steps. All words would be automatically considered
+safe up to $0$ steps and, for example, a read capability would be safe up to $n$
+steps if the world guarantees that the words at the corresponding addresses are
+safe up to $n - 1$ steps. We can then prove that the above read capability for
+address $a$ is safe up to any number of steps.
+
+\paragraph{Future worlds}
+So worlds are defined as a set of invariants on the memory, but what if we
+allocate fresh memory through malloc? We may want to establish new invariants
+for this freshly allocated memory and be sure that the adversary will also
+respect those (if we don't provide them with capabilities through which the new
+invariants can be broken). To accomodate this, we allow worlds to evolve, for
+example by adding additional invariants for freshly allocated memory. Formally,
+we define valid ways for a world $W$ to evolve into a new world $W'$ through the
+future-world relation $W' \future W$ and we ensure that the set of safe words in
+world $W$ must remain safe in any future world $W'$. Defining safety w.r.t. a
+notion of evolvable worlds makes our logical relation into a \emph{Kripke}
+logical relation.
+
+\paragraph{Invariants and Recursive Worlds}
+So worlds group a set of memory invariants, but how are they actually defined
+formally? We represent each invariant by a region $r \in \Regions$. We will see
+later that regions contain state machines to support a notion of evolvable
+invariants, but in every state, they also contain a predicate $H$ that defines
+which memory segments are acceptable in the current state of the invariant.
+Unfortunately, it is not enough to just take $H \in \powerset{\MemSegments}$,
+because sometimes the invariant may itself be world-dependent. For example, we
+may want to express invariants like ``the memory at address $50$ contains a
+value that is safe in the current world''. As explained, worlds may evolve, and
+the set of safe values may grow in future worlds, so then we need to take $H \in
+\Worlds \rightarrow \powerset{\MemSegments}$. We then end up with worlds
+containing regions and regions that are world-indexed, i.e. the set of worlds is
+recursively defined. Nevertheless, this problem is not essential: we will see
+how such a recursive definition can again be accomodated using techniques from
+the literature (essentially an advanced application of step-indexing).
+
+\paragraph{Local capabilities}
+A final thing we explain here is how our results take into account local
+capabilities and their special treatment by the hardware. Our idea here is to
+introduce a distinction between temporary and permanent invariants. At any point
+in a proof, we may decide to revoke temporary invariants. Global capabilities
+are then guaranteed to remain safe, but local ones aren't. Concretely, we will
+perform such revocations when we need to break the validity of capabilities that
+we previously provided to the adversary as a local capability (for example,
+return or stack pointers from a previous invocation). In such situations, the
+hardware guarantees that global capabilities (which cannot have write-local
+permission because of a central assumption\dominique{highlight earlier}) cannot
+have been used to store local capabilities and therefore remain safe even after
+revoking temporary invariants. Formally, this leads to two distinct future world
+relations: we say that $W' \futurewk W$ if a future world $W'$ preserves
+temporary invariants and $W' \futurestr W$ if it doesn't. Interestingly, this
+is a variant of a notion of public/private future worlds that has been previously
+used in the literature (see Section~\ref{sec:discussion}).
 %
@@
 % \lau{this may not be an important enough point to include it.}
-A world is a finite map from region names, modeled as natural numbers, to regions.
+Formally, a world is a finite map from region names, modeled as natural numbers, to regions.
 We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.

### 2017-06-01 popl-2018/paper.tex
Add a todo note...
Added comment about FTLR to answer Dominique's question.

@@
 notice that the adversary capability is an enter capability, which the
-Fundamental Theorem says nothing about.\dominique{why doesn't it?}  Luckily the enter capability becomes
+Fundamental Theorem says nothing about.\dominique{why doesn't it?} \lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
+The reason being that the copmutation would fail immediately and the condition in the observation relation would immediately be true (as it requires the computation to terminate and not fail).
+It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}  Luckily the enter capability becomes
 $\exec$ after the jump and then the Fundamental Theorem applies.
@@
 
+\dominique{Should we also discuss the fact that we only have a semantic model
+  but not a logic?}
+
 \section{Related work}

### 2017-06-06 popl-2018/paper.tex
updated format, notice we have fewer extra pages now
made a pass over sec. 4.1

@@
 %% For double-blind review submission
-\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
+\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
 %% For single-blind review submission
@@
 on the capabilities it has access to. Specifically, it does not matter which
-instructions are actually executed, i.e. the bound also applies to untrusted
+instructions are actually executed, i.e., the bound also applies to untrusted
 adversary code that has not been inspected or modified in any way.
@@
 authority bounds we consider are related to memory, but in a form that is more
-fine-grained than standard read/write authority: code's authority can be bounded
+fine-grained than standard read/write authority: a piece of code's authority can be bounded
 by arbitrary memory invariants that it is required to respect. Specifically, we
-will define worlds $W \in \Worlds$ that define a set of memory invariants and
-our results will express authority bounds on code as safety with respect to such
-a world, i.e. the fact that the code respects the invariants registered in
+will define worlds $W \in \Worlds$, which describe a set of memory invariants, and
+our results will express authority bounds on code as \emph{safety with respect to such
+a world,} i.e., the fact that the code respects the invariants registered in
 the world.
@@
 So, let's say that we have a world $W$ expressing that the memory must contain
-value $42$ at address $0$ must remain $42$, may contain arbitrary values at
-address $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$ and
+value $42$ at address $0$, may contain arbitrary values at
+addresses $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$, and
 an integer at address $100$ that may only increase over time\footnote{Indeed, we
-  will allow a notion of \emph{evolvable} invariants or protocols that can
+  will allow a notion of \emph{evolvable} invariants, aka \emph{protocols}, that can
   express such a temporal property.}. Our main theorem will state that if the
 current register file only contains safe words (numbers or capabilities which
-preserve the invariants in $W$ under any interaction), then the execution will
+preserve the invariants in $W$ under any interaction), then an execution will
 necessarily also preserve the invariants (irrespective of the instructions being
@@
 To make this more precise, we need to define the set $\stdvr(W) \in
-\powerset{\Words}$ of words that are safe w.r.t. $W$. Essentially, the set
+\powerset{\Words}$ of words that are safe w.r.t.\ $W$. Essentially, the set
 should only include words that preserve $W$'s invariants under any interaction,
@@
 More generally, we want to define that a read capability for memory range
-$[b,e]$ is safe if the world guarantees that the words at those addresses are
+$[b,e]$ is safe, if the world guarantees that the words at those addresses are
 themselves safe. However, this definition is cyclic: what if the world
@@
 $a$ is safe iff the same read capability for address $a$ is safe. This form of
-cyclic reasoning is related to similar problems in languages with recursive
+cyclic reasoning is related to similar challenges in languages with recursive
 types or higher-order ML-style references, and a standard solution is
-step-indexing. Essentially, the cycle is broken by defining safety up to a
-certain number of interaction steps. All words would be automatically considered
-safe up to $0$ steps and, for example, a read capability would be safe up to $n$
+to use step-indexing: Essentially, the cycle is broken by defining safety up to a
+certain number of interaction steps. All words willl be considered
+safe up to $0$ steps (since if there no interactions, nothing unsafe
+can happen), and, for example, a read capability will be safe up to $n$
 steps if the world guarantees that the words at the corresponding addresses are
@@
 example by adding additional invariants for freshly allocated memory. Formally,
-we define valid ways for a world $W$ to evolve into a new world $W'$ through the
+we define valid ways for a world $W$ to evolve into a new world $W'$ through a
 future-world relation $W' \future W$ and we ensure that the set of safe words in
-world $W$ must remain safe in any future world $W'$. Defining safety w.r.t. a
+world $W$ must remain safe in any future world $W'$. Defining safety w.r.t.\ a
 notion of evolvable worlds makes our logical relation into a \emph{Kripke}
@@
 \paragraph{Invariants and Recursive Worlds}
-So worlds group a set of memory invariants, but how are they actually defined
-formally? We represent each invariant by a region $r \in \Regions$. We will see
-later that regions contain state machines to support a notion of evolvable
-invariants, but in every state, they also contain a predicate $H$ that defines
-which memory segments are acceptable in the current state of the invariant.
-Unfortunately, it is not enough to just take $H \in \powerset{\MemSegments}$,
-because sometimes the invariant may itself be world-dependent. For example, we
-may want to express invariants like ``the memory at address $50$ contains a
-value that is safe in the current world''. As explained, worlds may evolve, and
-the set of safe values may grow in future worlds, so then we need to take $H \in
-\Worlds \rightarrow \powerset{\MemSegments}$. We then end up with worlds
-containing regions and regions that are world-indexed, i.e. the set of worlds is
-recursively defined. Nevertheless, this problem is not essential: we will see
-how such a recursive definition can again be accomodated using techniques from
-the literature (essentially an advanced application of step-indexing).
+So worlds group a set of memory invariants, but how are they actually
+defined formally? We represent each invariant by a region
+$r \in \Regions$. We will see later that regions contain state
+machines to support a notion of evolvable invariants, but in every
+state, they also contain a predicate $H$ that defines which memory
+segments are acceptable in the current state of the invariant.
+Unfortunately, it is not enough to just take
+$H \in \powerset{\MemSegments}$, because sometimes the invariant may
+itself be world-dependent. For example, we may want to express
+invariants like ``the memory at address $50$ contains a value that is
+safe in the current world''. As explained, worlds may evolve, and the
+set of safe values may grow in future worlds, and therefore we need to
+index $H$ over worlds, i.e., take
+$H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
+with worlds containing regions and regions that are world-indexed,
+i.e., the set of worlds must be recursively defined. We will see how such a recursive
+definition can be accomodated using techniques from the
+literature (essentially an advanced application of step-indexing).
 

### 2017-06-07 popl-2018/paper.tex
typos

@@
 proof should depend on a formal expression of the guarantees provided by the
-capability machine, including the specific guarantees for local capbalities.
+capability machine, including the specific guarantees for local capabilities.
 
@@
 to use step-indexing: Essentially, the cycle is broken by defining safety up to a
-certain number of interaction steps. All words willl be considered
+certain number of interaction steps. All words will be considered
 safe up to $0$ steps (since if there no interactions, nothing unsafe
@@
 respect those (if we don't provide them with capabilities through which the new
-invariants can be broken). To accomodate this, we allow worlds to evolve, for
+invariants can be broken). To accommodate this, we allow worlds to evolve, for
 example by adding additional invariants for freshly allocated memory. Formally,
@@
 i.e., the set of worlds must be recursively defined. We will see how such a recursive
-definition can be accomodated using techniques from the
+definition can be accommodated using techniques from the
 literature (essentially an advanced application of step-indexing).
@@
 Fundamental Theorem says nothing about.\dominique{why doesn't it?} \lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
-The reason being that the copmutation would fail immediately and the condition in the observation relation would immediately be true (as it requires the computation to terminate and not fail).
+The reason being that the computation would fail immediately and the condition in the observation relation would immediately be true (as it requires the computation to terminate and not fail).
 It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}  Luckily the enter capability becomes

### 2017-06-07 popl-2018/paper.tex
updated figures to better fit with the new format

@@
 \newcommand{\decodePerm}[1]{\plainfun{decodePerm}{#1}}
-\newcommand{\updatePcPerm}[1]{\plainfun{updatePcPerm}{#1}}
+\newcommand{\updatePcPerm}[1]{\plainfun{updPcPerm}{#1}}
 
@@
 \newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
-\newcommand{\stdUpdatePc}[1]{\plainfun{updatePc}{#1}}
+\newcommand{\stdUpdatePc}[1]{\plainfun{updPc}{#1}}
 
@@
 \newcommand{\Mems}{\Heaps}
-%\newcommand{\HeapSegments}{\plaindom{MemSegment}}
 \newcommand{\HeapSegments}{\plaindom{MemSeg}}
@@
 \begin{figure}
-\[
-  \begin{array}{rlrcl p{2cm}rlrcl}
-    \addr   &\in& \Addrs &\isdef& \nats & & r &\in& \RegName&\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
-    w &\in&\Words &\isdef& \ints + \Caps & & \reg &\in& \Regs  &\isdef& \RegName \rightarrow \Words\\
-    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m&\in& \Heaps &\isdef& \Addrs \rightarrow \Words \\
-    & & & &\exec\mid \entry\mid \rwx\mid \rwlx & & \Phi &\in& \ExecConfs  &\isdef& \Regs \times \Heaps \\
-    \gl&\in&\Globals & ::= &\glob \mid \local & & \ms &\in& \MemSegments &::= & \Addrs \parfun \Words \\
-    &&\Confs &::=& \multicolumn{5}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
-     && \Caps  &::= & \multicolumn{5}{l}{\left\{
+\[\arraycolsep=1.4pt
+  \begin{array}{rlrcl p{5cm}rlrcl}
+    \addr   &\in& \Addrs &\isdef& \nats &\phantom{makesomero}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
+    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words\\
+    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
+    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
+    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
+    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
+     && \Caps  &::= & \multicolumn{6}{l}{\left\{
                          ((\perm,\gl),\start,\addrend,\addr)\mid
@@
                  \minus{\lv}{\rv}{\rv} \mid 
-                 \lea{\lv}{\rv} \mid 
-                 \restricttwo{\lv}{\rv} \mid \\
-           &   & \subseg{\lv}{\rv}{\rv} \mid  
+                 \lea{\lv}{\rv} \mid \\
+          &   &  \restricttwo{\lv}{\rv} \mid
+                 \subseg{\lv}{\rv}{\rv} \mid  
                  \isptr{\lv}{\rv} \mid 
@@
                  \getl{\lv}{\lv} \mid \getb{\lv}{\lv} \mid
-                 \gete{\lv}{\lv} \mid
-                 \geta{\lv}{\lv} \mid 
+                 \gete{\lv}{\lv} \mid \\
+          &   &  \geta{\lv}{\lv} \mid 
                  \fail \mid
@@
       \begin{array}[t]{l}
-\text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$}\\ \text{\;\;and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }
+        \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$}\\ 
+        \text{\;\;and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }
       \end{array}
@@
                            \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
-                             \text{if $\memreg(\pcreg) = \stdcap$ and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}\\
+                           \begin{array}[t]{l}
+                             \text{if $\memreg(\pcreg) = \stdcap$}\\
+                             \text{\;\;and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
+                           \end{array}
+\\
                              \failed & \text{otherwise}
@@
   \end{align*}
-  \begin{tabular}{|c|p{3.9cm}|p{8.3cm}|}
+  \begin{tabular}{|c|p{3.4cm}|p{7.3cm}|}
     \hline
@@
     \hline
-    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and $w = \memheap(\addr)$ \newline and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
+    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and  $w = \memheap(\addr)$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
     \hline
@@
     \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$\newline otherwise $\var{newPc} = \memreg(r)$\\
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
     \hline
@@
       & \observations \; : \;  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-      & \observations (W) \defeq \\
-      & \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
+      & \observations (W) \defeq 
+        \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
         \begin{aligned}
-          & \forall \ms_f, \heap', i \leq n \ldotp \\
-          & (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
+          & \forall \ms_f, \heap', i \leq n \ldotp  (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
           & \;\; \Rightarrow
@@
             & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
-            & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \\
-            & \;\; \heapSat[\hs']{n-i}{W'}
+            & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \heapSat[\hs']{n-i}{W'}
           \end{aligned}
@@
       &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
-      &\stdvr(W)\defeq  \\
-      &\;\;\begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \}
-        \union \\
-        & \{ \npair{\stdcap[(\noperm,\gl)] } \}
+      &\stdvr(W)\defeq 
+        \begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \} \union \{ \npair{\stdcap[(\noperm,\gl)] } \}
         \union \\
         & \left\{
-          \begin{multlined}
-            \npair{\stdcap[(\readonly,\gl)] } \mid \\
-            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-          \end{multlined} \right\}
+            \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+          \right\}
         \union \\
         & \left\{
-          \begin{aligned}
-            &\npair{\stdcap[(\readwrite,\gl)] } \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             &\quad\npair{(\start,\addrend)} \in
-            \writeCond{}(\iota^\nwl,\gl)(W)
-          \end{aligned} \right\} \union \\
+            \npair{\stdcap[(\readwrite,\gl)] } \; \middle| \;
+            \begin{array}{l}
+              \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+              \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
+            \end{array}
+          \right\} \union \\
         & \left\{
-          \begin{aligned}
-            &\npair{\stdcap[(\readwritel,\gl)] } \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             &\quad\npair{(\start,\addrend)} \in
-            \writeCond{}(\iota^\pwl,\gl)(W)
-          \end{aligned} \right\}
+            \npair{\stdcap[(\readwritel,\gl)] } \;\middle| \;
+            \begin{array}{l}
+             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+             \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
+            \end{array}
+           \right\}
         \union \\
         & \left\{
-          \begin{aligned}
-            &\npair{\stdcap[(\exec,\gl)]} \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             &\quad\npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-          \end{aligned} \right\}
+          \npair{\stdcap[(\exec,\gl)]} \; \middle| \;
+          \begin{array}{l}
+             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+             \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{array}
+           \right\}
         \union \\
         & \left\{
-          \begin{multlined}
-            \npair{\stdcap[(\entry,\gl)]} \mid \\
+            \npair{\stdcap[(\entry,\gl)]} \; \middle| \;
             \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)
-          \end{multlined} \right\}
+          \right\}
         \union \\
         & \left\{
-          \begin{aligned}
-            &\npair{\stdcap[(\rwx,\gl)]} \mid \\
-            &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-            &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
-            &\quad\npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-          \end{aligned} \right\}
+          \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
+          \begin{array}{l}
+            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+            \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
+            \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{array}
+          \right\}
         \union \\
         & \left\{
-          \begin{aligned}
-            &\npair{\stdcap[(\rwlx,\gl)]} \mid \\
-             &\quad\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             &\quad\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
-             &\quad\npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in
+            \npair{\stdcap[(\rwlx,\gl)]} \; \middle| \;
+            \begin{array}{l}
+             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+             \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
+             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in
             \execCond{}(\gl)(W)
-          \end{aligned} \right\}
+            \end{array}
+          \right\}
       \end{aligned}
@@
 %  \end{withmathindent}
-  \caption{Logical Relation}
+  \caption{Logical relation}
   \label{fig:logrel}            
@@
   \begin{align*}
-  & \readCond{}(\gl)(W) =  \\
-  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
-    \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
-    \end{aligned} \right\}\\
-  & \writeCond{}(\iota,\gl)(W) =  \\
-  & \left\{
+  & \readCond{}(\gl)(W) =  \left\{ \npair{(\start,\addrend)} \; \middle| \;
+       \exists r \in \var{localityReg}(g,W) \ldotp 
+       \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
+    \right\}\\
+  & \writeCond{}(\iota,\gl)(W) =  \left\{
     \npair{(\start,\addrend)}
@@
     \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & \;\; W(r) \text{ is address-stratified and} \\
+      & \exists r \in \var{localityReg}(g,W) \ldotp W(r) \text{ is address-stratified and} \\
       & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
     \end{aligned} \right\}\\
-  & \execCond{}(\gl)(W) = \\
-  & \left\{
+  & \execCond{}(\gl)(W) = 
+    \left\{
     \begin{multlined}
@@
     \end{aligned} \right\} \\
-  & \entryCond{}(\gl)(W) = \\
-  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
-    \begin{aligned}
- &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
-      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
-    \end{aligned} \right\} \\
+  & \entryCond{}(\gl)(W) = 
+    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
+    \forall n' < n \ldotp \forall W' \future W \ldotp \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
+    \right\} \\
   & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
@@
       \begin{aligned}[t]
-        &\perm = \rwx \text{ and } \\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
+        &\perm = \rwx \text{ and } \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
         &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
@@
       \begin{aligned}[t]
-        &\perm = \rwlx \text{ and }\\
-        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
+        &\perm = \rwlx \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
         &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
@@
     c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
-    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
-      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
+    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
     \end{array}

### 2017-06-12 popl-2018/paper.tex
Fix typo.

@@
 values. They are a kind of values that can be provided to other code for
-immediate use without allowing it to be stored in a closure or reference for
+immediate use without allowing them to be stored in a closure or reference for
 later use. We believe reasoning about such values will require techniques similar 

### 2017-06-12 popl-2018/paper.tex
Changed enter permission symbol to have a bar over it to avoid confusion with variable for end of interval.
Added a small explanation of variables in Thm 1 and gave some intuition relating to previous text (reviewer 3 found that this was missing).

--- 2017-06-07 popl-2018/paper.tex
@@
 \newcommand{\exec}{\plainperm{rx}}
-\newcommand{\entry}{\plainperm{e}}
+\newcommand{\entry}{\bar{\plainperm{e}}}
 \newcommand{\rwx}{\plainperm{rwx}}
@@
 \end{theorem}
-We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
-the worlds:
+In the above theorem, $\States \times \Rels$ corresponds to the
+aforementioned state transition system where $\Rel$ is a set of pairs
+of relations corresponding to the public and private transition, and
+$\States$ is a sufficiently large set of states containing at least the
+states we use. The last part of the temporary and permanent regions is
+a state interpretation function that determines what memory segments
+the region permits in each state of its states. The different
+monotonicity requirements in the two interpretation functions reflects
+how permanent regions rely only on permanent protocols whereas
+temporary regions can rely on both temporary and permanent protocols.
+$\UPred{\HeapSegments}$ is a step-indexed, downwards closed predicate
+on memory segments.
+
+With the recursive domain equation solved, we use the $\Regions$ from
+Theorem~\ref{thm:world-existence} to define the worlds:
 \[

### 2017-06-13 popl-2018/paper.tex
Add a discussion of semantic vs. syntactic properties.

--- 2017-06-12 popl-2018/paper.tex
@@
 also global. \\\\
+
+\noindent \textbf{Reasoning about capability machine programs} \emph{Semantic,
+  but not syntactic properties} The logical relation defined in
+Section~\ref{sec:logical-relation} allows us to reason about capability machine
+programs. A limitation w.r.t. previous work is that the logical relation is
+tailored exclusively towards semantic properties, not syntactic ones.
+
+Imagine, for example, that we invoke a block of adversary code in such a way
+that it only ever receives capabilities within a specific range of memory. After
+the code returns, we may try to prove that any capabilities passed back to us in
+the registers are still confined to that range of memory. The property of
+falling in a certain address range is syntactic in the sense that it talks about
+the specific implementation of a higher-order value rather than its behavior,
+like the invariants that are required/preserved when we use it.
+
+Such syntactic properties are hard to prove in our system. For the example
+cited, it would be easy to conclude that the returned values are in the value
+relation (see Figure~\ref{fig:logrel}). This gives us a lot of semantic
+information, like conditions under which they are safe to use and invariants
+that will be preserved when we do, but it does not tell us much about the
+address of the capability.
+
+To make this very concrete, consider how capabilities with permission $\noperm$
+are always in the value relation, irrespective of their properties.
+Semantically, this makes perfect sense, since they are always safe since they
+cannot be used for anything. However, it also means we do not get syntactic
+information about them.
+
+For our purposes, this restriction is unproblematic, since we are only
+interested in proving semantic properties (e.g., an assertion will never fail).
+However, in other situations, we may be interested in proving more syntactic
+properties like the more syntactic properties, that are often considered in
+object capability literature like confinement, no authority amplification etc.
+Although such properties are typically more restrictive and less easy to use for
+reasoning, \cite{Devriese:2016ObjCap} have demonstrated how a logical relation
+much like ours can be adapted to also support them, by quantifying the logical
+relation over a custom interpretation of effectful computations and the type of
+references. We expect their solution can be readily adapted to our setting,
+modulo some details (like the fact that we do not just have read-write
+capabilities, but also other types).
+
+% \emph{No program logic} TODO: discuss this? 
+\\\\
+
 \noindent\textbf{Logical relation}\\

### 2017-06-13 popl-2018/paper.tex
added a bit of text pointing out macros make examples small and readable

@@
 \subsubsection{Future worlds}
-The future world relations model how memory may evolve over time. The
-\emph{public future world} $W' \futurewk W$ requires that $\dom(W') \supseteq
+\label{subsec:future-worlds} 
+The future world relations model how memory may evolve over time. 
+The \emph{public future world} $W' \futurewk W$ requires that $\dom(W') \supseteq
 \dom(W)$ and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$. That is, in a
@@
 
-Notice that the public future region relation is a subset of the
-private future region relation.
+% Notice that the public future region relation is a subset of the
+% private future region relation.
 
@@
 \label{sec:examples}
-In this section, we demonstrate how our formalization of capability safety
-allows us to prove local-state encapsulation and control-flow correctness
-properties for challenging program examples. The security measures of
-Section~\ref{sec:stack-and-return-pointer} are deployed to ensure these
-properties. Since we are dealing with assembly language, there are many details
-to the formal treatment, and therefore we necessarily have to omit some details
-in the lemmas we state here. The interested reader can find all the technical
-details in the technical appendix~\citep{technical_appendix}.
+In this section, we demonstrate how our formalization of capability
+safety allows us to prove local-state encapsulation and control-flow
+correctness properties for challenging program examples. The security
+measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
+ensure these properties. Since we are dealing with assembly language,
+there are many details to the formal treatment, and therefore we
+necessarily have to omit some details in the lemmas we state here. The
+examples may look deceivingly short, but it is because they use the
+previously described macro instructions, which expands to multiple basic instructions. The examples would be unintelligible without the macros. The interested
+reader can find all the technical details in the technical
+appendix~\citep{technical_appendix}.
 
@@
 Although such properties are typically more restrictive and less easy to use for
-reasoning, \cite{Devriese:2016ObjCap} have demonstrated how a logical relation
+reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation
 much like ours can be adapted to also support them, by quantifying the logical
@@
 modulo some details (like the fact that we do not just have read-write
-capabilities, but also other types).
+capabilities, but also other types).\\\\
 
 % \emph{No program logic} TODO: discuss this?
-\\\\
 

### 2017-06-14 popl-2018/paper.tex
Comment out boilerplate acks.

@@
 %% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
+% \begin{acks}                            %% acks environment is optional
+%                                         %% contents suppressed with 'anonymous'
+%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+%   %% acknowledge financial support and will be used by metadata
+%   %% extraction tools.
+%   This material is based upon work supported by the
+%   \grantsponsor{GS100000001}{National Science
+%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+%   conclusions or recommendations expressed in this material are those
+%   of the author and do not necessarily reflect the views of the
+%   National Science Foundation.
+% \end{acks}
 

### 2017-06-14 popl-2018/paper.tex
Fix compilation further.

@@
 capabilities, but also other types).
-
 % \emph{No program logic} TODO: discuss this?
-\\\\
 

### 2017-06-14 popl-2018/paper.tex
Added reasonable compiler provides protection (no detailed explanation).

@@
 stack overwriting activation records' contents). Luckily, such protection is
-easy to provide thanks to the capability machine's fine-grained memory
-protection\dominique{explain this point better}.
+easy for a reasonable compiler to provide thanks to the capability machine's
+fine-grained memory protection\dominique{explain this point better}.
 
@@
 examples may look deceivingly short, but it is because they use the
-previously described macro instructions, which expands to multiple basic instructions. The examples would be unintelligible without the macros. The interested
-reader can find all the technical details in the technical
-appendix~\citep{technical_appendix}.
+previously described macro instructions, which expands to multiple
+basic instructions. The examples would be unintelligible without the
+macros. The interested reader can find all the technical details in
+the technical appendix~\citep{technical_appendix}.
 
@@
 that the configuration just after \texttt{\footnotesize{scall}} is done cleaning
-up is safe. We use the Fundamental Theorem to reason about the unknown code, but
+up is safe. We use the Fundamental Theorem to reason about the unknown adversarial code, but
 notice that the adversary capability is an enter capability, which the
-Fundamental Theorem says nothing about.\dominique{why doesn't it?} \lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
+Fundamental Theorem says nothing about.
+\dominique{why doesn't it?}
+\lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
 The reason being that the computation would fail immediately and the condition in the observation relation would immediately be true (as it requires the computation to terminate and not fail).
-It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}  Luckily the enter capability becomes
+It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}
+Luckily the enter capability becomes
 $\exec$ after the jump and then the Fundamental Theorem applies.

### 2017-06-14 popl-2018/paper.tex
Expand the discussion of local capabilities in the introductory section of the LR.

--- 2017-06-14 popl-2018/paper.tex
@@
 A final thing we explain here is how our results take into account local
-capabilities and their special treatment by the hardware. Our idea here is to
-introduce a distinction between temporary and permanent invariants. At any point
-in a proof, we may decide to revoke temporary invariants. Global capabilities
-are then guaranteed to remain safe, but local ones aren't. Concretely, we will
-perform such revocations when we need to break the validity of capabilities that
-we previously provided to the adversary as a local capability (for example,
-return or stack pointers from a previous invocation). In such situations, the
-hardware guarantees that global capabilities (which cannot have write-local
-permission because of a central assumption\dominique{highlight earlier}) cannot
-have been used to store local capabilities and therefore remain safe even after
-revoking temporary invariants. Formally, this leads to two distinct future world
-relations: we say that $W' \futurewk W$ if a future world $W'$ preserves
-temporary invariants and $W' \futurestr W$ if it doesn't. Interestingly, this
-is a variant of a notion of public/private future worlds that has been previously
-used in the literature (see Section~\ref{sec:discussion}).
-%
+capabilities and their special treatment by the hardware. Normally, when we
+invoke an untrusted piece of code and provide it with certain global
+capabilities, it may have stored those capabilities in memory and we will only
+be able to reinvoke the code if we can guarantee that those values are still
+valid. Formally, worlds will represent the invariants that the global
+capabilities' safety relies on and reinvocation will only be possible in future
+worlds.
+
+However, if we provide the adversary with local capabilities in that first
+invocation, then the situation is a bit different. Since the adversary has no
+way to store local capabilities, we can allow the second invocation to happen in
+\emph{private} future worlds ($W' \futurestr W$), in which global capabilities
+remain safe, but local ones don't. This private future world relation is more
+liberal than the public one ($W' \futurewk W$, in which \emph{all} safe
+capabilities remain safe). In particular, worlds may contain \emph{temporary}
+regions, representing invariants that only local capabilties may rely on for
+their safety, and which may be revoked (disabled) in private future worlds.
+
+Interestingly, this idea is a variant of a notion of public/private future
+worlds that has been previously used in the literature (see
+Section~\ref{sec:discussion}). However, temporary regions are new in our setting
+and there is an interesting interplay with the recursiveness of the worlds: for
+a temporary region, the predicate $H \in \Worlds \rightarrow
+\powerset{\MemSegments}$ (which defines which memory segments are safe in the
+current state of the world) is only required to be monotonous w.r.t.\ public
+future worlds (i.e.\ safe memory segments remain safe in public future worlds),
+but for permanent regions, it must be monotonous w.r.t.\ private future worlds.
+As a consequence, the memory for a permanent region may not contain local
+capabilities (as their safety would be broken in private future worlds), which
+in turn implies that only local capabilities may have store-local permission (a
+general sanity requirement when using local capabilities\footnote{In fact, local
+  capabilities become useless as soon as the adversary has access to a single
+  global, store-local capability.}).
 
@@
 
-\noindent \textbf{Reasoning about capability machine programs} \emph{Semantic,
-  but not syntactic properties} The logical relation defined in
+\noindent \textbf{Reasoning about capability machine programs}\\
+\emph{Semantic, but not syntactic properties} The logical relation defined in
 Section~\ref{sec:logical-relation} allows us to reason about capability machine
@@
 modulo some details (like the fact that we do not just have read-write
-capabilities, but also other types).
+capabilities, but also other types).\\\\
 % \emph{No program logic} TODO: discuss this?

### 2017-06-14 popl-2018/paper.tex
Question about calling convention.
Added a mention that local capabilities can be used for more that CC. Needs more work, maybe add an example

--- 2017-06-14 popl-2018/paper.tex
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine the adversary stores their stack pointer, invokes trusted
-code that uses part of the stack to store private data and then invokes the
-adversary again with a stack pointer restricted to exclude the part containing
-the private data. If the adversary had a way of keeping hold of their old stack
-pointer, it could access the private data stored there by the trusted code and
-break local-state encapsulation.
+pointers. Imagine the adversary stores their stack pointer, invokes trusted code
+that uses part of the stack to store private data and then invokes the adversary
+again with a stack pointer restricted to exclude the part containing the private
+data. If the adversary had a way of keeping hold of their old stack pointer, it
+could access the private data stored there by the trusted code and break
+local-state encapsulation.  
+\lau{Do we say anything about arguments to the callback?
+Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary.
+Can we even to any extend allow local arguments that we pass on from an unknown source?}
 
@@
 \end{description}
+
+\textbf{A word on local capabilities} Local capabilities are an essential part
+of the calling convention. The use of local capabilities is, however, not
+limitted to the stack and return pointer.
+
 \textbf{Reusable macro instructions} We define a number of reusable

### 2017-06-14 popl-2018/paper.tex
Changed all permissions to use small caps - no bar
typo

@@
 % Permissions
-\newcommand{\plainperm}[1]{\mathrm{#1}}
+\newcommand{\plainperm}[1]{\textsc{#1}}
 \newcommand{\noperm}{\plainperm{o}}
@@
 \newcommand{\exec}{\plainperm{rx}}
-\newcommand{\entry}{\bar{\plainperm{e}}}
+\newcommand{\entry}{\plainperm{e}}
 \newcommand{\rwx}{\plainperm{rwx}}
@@
 % Global/local
-\newcommand{\local}{\plainperm{local}}
-\newcommand{\glob}{\plainperm{global}}
+\newcommand{\plainlocality}[1]{\mathrm{#1}}
+\newcommand{\local}{\plainlocality{local}}
+\newcommand{\glob}{\plainlocality{global}}
 
@@
 In the above theorem, $\States \times \Rels$ corresponds to the
-aforementioned state transition system where $\Rel$ is a set of pairs
+aforementioned state transition system where $\Rels$ is a set of pairs
 of relations corresponding to the public and private transition, and

### 2017-06-14 popl-2018/paper.tex
added adversary unknown piece of code

@@
 efficient, does not rely on a trusted central stack manager, and supports
-higher-order interfaces to an adversary. In this section, we explain this
+higher-order interfaces to an adversary where an adversary is just some unknown piece of code. In this section, we explain this
 convention's high-level approach, the security measures to be taken in a number

### 2017-06-14 popl-2018/paper.tex
updated contributions to specifically state it is a novel LR.

@@
   support: an efficient instruction for clearing a range of memory.
-\item Prove a theorem that is the most general and powerful formulation of the
-  guarantees offered by a capability machine (a form of capability
+\item A novel Kripke logical relation for reasoning about programs on the
+  capability machine. The fundamental theorem of logical relations here gives
+  the most general and powerful formulation of the guarantees offered by a
+  capability machine (a form of capability
   safety~\citet{Devriese:2016ObjCap,Maffeis2010OC}), including the specific

### 2017-06-15 popl-2018/paper.tex
Fix the problem pointed out by Lau in my introduction to public/private future worlds.

--- 2017-06-14 popl-2018/paper.tex
@@
 % Permissions
-\newcommand{\plainperm}[1]{\textsc{#1}}
+\newcommand{\plainperm}[1]{\mathrm{#1}}
 \newcommand{\noperm}{\plainperm{o}}
@@
 % Global/local
-\newcommand{\plainlocality}[1]{\mathrm{#1}}
-\newcommand{\local}{\plainlocality{local}}
-\newcommand{\glob}{\plainlocality{global}}
+\newcommand{\local}{\plainperm{local}}
+\newcommand{\glob}{\plainperm{global}}
 
@@
 efficient, does not rely on a trusted central stack manager, and supports
-higher-order interfaces to an adversary where an adversary is just some unknown piece of code. In this section, we explain this
+higher-order interfaces to an adversary. In this section, we explain this
 convention's high-level approach, the security measures to be taken in a number
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine the adversary stores their stack pointer, invokes trusted code
-that uses part of the stack to store private data and then invokes the adversary
-again with a stack pointer restricted to exclude the part containing the private
-data. If the adversary had a way of keeping hold of their old stack pointer, it
-could access the private data stored there by the trusted code and break
-local-state encapsulation.  
-\lau{Do we say anything about arguments to the callback?
-Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary.
-Can we even to any extend allow local arguments that we pass on from an unknown source?}
+pointers. Imagine the adversary stores their stack pointer, invokes trusted
+code that uses part of the stack to store private data and then invokes the
+adversary again with a stack pointer restricted to exclude the part containing
+the private data. If the adversary had a way of keeping hold of their old stack
+pointer, it could access the private data stored there by the trusted code and
+break local-state encapsulation.
 
@@
 stack overwriting activation records' contents). Luckily, such protection is
-easy for a reasonable compiler to provide thanks to the capability machine's
-fine-grained memory protection\dominique{explain this point better}.
+easy to provide thanks to the capability machine's fine-grained memory
+protection\dominique{explain this point better}.
 
@@
 \end{description}
-
-\textbf{A word on local capabilities} Local capabilities are an essential part
-of the calling convention. The use of local capabilities is, however, not
-limitted to the stack and return pointer.
-
 \textbf{Reusable macro instructions} We define a number of reusable
@@
 be able to reinvoke the code if we can guarantee that those values are still
-valid. Formally, worlds will represent the invariants that the global
-capabilities' safety relies on and reinvocation will only be possible in future
-worlds.
+valid. Formally, worlds represent the invariants that the global capabilities'
+safety relies on and the reinvocation is only safe in future worlds, where the
+invariants are respected.
 
 However, if we provide the adversary with local capabilities in that first
-invocation, then the situation is a bit different. Since the adversary has no
-way to store local capabilities, we can allow the second invocation to happen in
-\emph{private} future worlds ($W' \futurestr W$), in which global capabilities
-remain safe, but local ones don't. This private future world relation is more
-liberal than the public one ($W' \futurewk W$, in which \emph{all} safe
-capabilities remain safe). In particular, worlds may contain \emph{temporary}
-regions, representing invariants that only local capabilties may rely on for
-their safety, and which may be revoked (disabled) in private future worlds.
+invocation, then the situation is a bit different. The adversary has no way to
+store these local capabilities, so if we make sure that there are also no local
+capabilities in the register file for the second invocation (including the
+capability being invoked), then the adversary cannot use them any more.%
+\footnote{We are ignoring store-local capabilities in this discussion. If the
+  adversary does have access to store-local capabilities in the first and second
+  invocation, then the memory they address must entirely be cleared before the
+  second invocation in order for the reinvocation to remain safe.} Therefore, we
+can allow the second invocation to happen in \emph{private} future worlds ($W'
+\futurestr W$), in which global safe capabilities remain safe, but local ones
+don't. This private future world relation is more liberal than the standard,
+public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
+safe). Concretely, worlds may contain \emph{temporary} regions, representing
+invariants that only local capabilties may rely on for their safety, and which
+may be revoked (disabled) in private future worlds.
 
@@
 In the above theorem, $\States \times \Rels$ corresponds to the
-aforementioned state transition system where $\Rels$ is a set of pairs
+aforementioned state transition system where $\Rel$ is a set of pairs
 of relations corresponding to the public and private transition, and
@@
 examples may look deceivingly short, but it is because they use the
-previously described macro instructions, which expands to multiple
-basic instructions. The examples would be unintelligible without the
-macros. The interested reader can find all the technical details in
-the technical appendix~\citep{technical_appendix}.
+previously described macro instructions, which expands to multiple basic instructions. The examples would be unintelligible without the macros. The interested
+reader can find all the technical details in the technical
+appendix~\citep{technical_appendix}.
 
@@
 that the configuration just after \texttt{\footnotesize{scall}} is done cleaning
-up is safe. We use the Fundamental Theorem to reason about the unknown adversarial code, but
+up is safe. We use the Fundamental Theorem to reason about the unknown code, but
 notice that the adversary capability is an enter capability, which the
-Fundamental Theorem says nothing about.
-\dominique{why doesn't it?}
-\lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
+Fundamental Theorem says nothing about.\dominique{why doesn't it?} \lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
 The reason being that the computation would fail immediately and the condition in the observation relation would immediately be true (as it requires the computation to terminate and not fail).
-It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}
-Luckily the enter capability becomes
+It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}  Luckily the enter capability becomes
 $\exec$ after the jump and then the Fundamental Theorem applies.

### 2017-06-15 popl-2018/paper.tex
store-local -> write-local

@@
 capability being invoked), then the adversary cannot use them any more.%
-\footnote{We are ignoring store-local capabilities in this discussion. If the
-  adversary does have access to store-local capabilities in the first and second
+\footnote{We are ignoring write-local capabilities in this discussion. If the
+  adversary does have access to write-local capabilities in the first and second
   invocation, then the memory they address must entirely be cleared before the
@@
 capabilities (as their safety would be broken in private future worlds), which
-in turn implies that only local capabilities may have store-local permission (a
+in turn implies that only local capabilities may have write-local permission (a
 general sanity requirement when using local capabilities\footnote{In fact, local
   capabilities become useless as soon as the adversary has access to a single
-  global, store-local capability.}).
+  global, write-local capability.}).
 

### 2017-06-15 popl-2018/paper.tex
Add a footnote about a certain simplification made for explanation purposes in the discussion about local capabilities/temporary regions/public future worlds.

@@
   capabilities become useless as soon as the adversary has access to a single
-  global, write-local capability.}).
+  global, write-local capability.})\footnote{For explanation purposes, this
+  discussion ignores certain ways to allow for local capabilities in a permanent
+  region, for example, by not requiring that they are valid or requiring that
+  they are local versions of valid global capabilities etc.}.
 

### 2017-06-15 popl-2018/paper.tex
rewrote a bit of Worlds subsection to fit better with new introduction

--- 2017-06-15 popl-2018/paper.tex
@@
 \subsection{Formalizing the guarantees of the capability machine}
+\label{sec:formalizing-guarantees}
 What differentiates a capability machine from a more standard assembly language
@@
 certain number of interaction steps. All words will be considered
-safe up to $0$ steps (since if there no interactions, nothing unsafe
+safe up to $0$ steps (since if there is no interaction, nothing unsafe
 can happen), and, for example, a read capability will be safe up to $n$
@@
 $H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
-with worlds containing regions and regions that are world-indexed,
+with worlds containing regions with world-indexed predicates,
 i.e., the set of worlds must be recursively defined. We will see how such a recursive
@@
 be able to reinvoke the code if we can guarantee that those values are still
-valid. Formally, worlds represent the invariants that the global capabilities'
-safety relies on and the reinvocation is only safe in future worlds, where the
-invariants are respected.
-
-However, if we provide the adversary with local capabilities in that first
-invocation, then the situation is a bit different. The adversary has no way to
-store these local capabilities, so if we make sure that there are also no local
-capabilities in the register file for the second invocation (including the
-capability being invoked), then the adversary cannot use them any more.%
-\footnote{We are ignoring store-local capabilities in this discussion. If the
-  adversary does have access to store-local capabilities in the first and second
-  invocation, then the memory they address must entirely be cleared before the
-  second invocation in order for the reinvocation to remain safe.} Therefore, we
-can allow the second invocation to happen in \emph{private} future worlds ($W'
-\futurestr W$), in which global safe capabilities remain safe, but local ones
-don't. This private future world relation is more liberal than the standard,
-public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
-safe). Concretely, worlds may contain \emph{temporary} regions, representing
-invariants that only local capabilties may rely on for their safety, and which
-may be revoked (disabled) in private future worlds.
+valid. Formally, worlds will represent the invariants that the global
+capabilities' safety relies on and reinvocation will only be possible in future
+worlds.
+
+However, if we provide the adversary with local capabilities in the first
+invocation, then the situation is a bit different. Since the adversary has no
+way to store local capabilities, we can allow the second invocation to happen in
+\emph{private} future worlds ($W' \futurestr W$), in which global capabilities
+remain safe, but local ones don't. This private future world relation is more
+liberal than the public one ($W' \futurewk W$, in which \emph{all} safe
+capabilities remain safe). In particular, worlds may contain \emph{temporary}
+regions, representing invariants that only local capabilties may rely on for
+their safety, and which may be revoked (disabled) in private future worlds.
 
@@
 % \lau{this may not be an important enough point to include it.}
-Formally, a world is a finite map from region names, modeled as natural numbers, to regions.
-We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
+
+As describe above, worlds are a collection of invariants. Formally, a world is a
+finite map from region names, modeled as natural numbers, to regions that each
+correspond to an invariant of part of the memory.  We have three types of
+regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
 %$\perma$, $\temp$, and $\revoked$. The
@@
 Each permanent and temporary region contains a state transition system, with
-public and private transitions, to describe a protocol for the region's memory,
-similarly to what has been used in logical relations for high-level
+public and private transitions, to describe how the invariants are allowed to
+change over time. In other words, they are protocols for the region's
+memory. These are similar to what has been used in logical relations for
+high-level
 languages~\citep{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
@@
 %\dominique{``end up being defined recursively'' does not sound very insightful}
-A region defines what its part of memory can legally look like. However,
-validity of the memory's contents may itself depend on the state of other parts
-of memory. To accommodate this, worlds need to be defined recursively: Worlds
-describe memory, which can contain capabilities, whose safety is
-world-dependent. Recursive worlds are a common requirement in Kripke models and
-the following lemma uses the method of
+% A region defines what its part of memory can legally look like. However,
+% validity of the memory's contents may itself depend on the state of other parts
+% of memory. To accommodate this, worlds need to be defined recursively: Worlds
+% describe memory, which can contain capabilities, whose safety is
+% world-dependent. 
+As described in Section~\ref{sec:formalizing-guarantees}, the safety of a world
+depends on the safety of its regions which in turn may depend on the safety of
+the values it contain which depend on the safety of a world. In other words, we
+need recursive worlds. Recursive worlds are a common requirement in Kripke
+models and the following lemma uses the method of
 \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for

### 2017-06-15 popl-2018/paper.tex

@@
 be able to reinvoke the code if we can guarantee that those values are still
-valid. Formally, worlds will represent the invariants that the global
-capabilities' safety relies on and reinvocation will only be possible in future
-worlds.
-
-However, if we provide the adversary with local capabilities in the first
-invocation, then the situation is a bit different. Since the adversary has no
-way to store local capabilities, we can allow the second invocation to happen in
-\emph{private} future worlds ($W' \futurestr W$), in which global capabilities
-remain safe, but local ones don't. This private future world relation is more
-liberal than the public one ($W' \futurewk W$, in which \emph{all} safe
-capabilities remain safe). In particular, worlds may contain \emph{temporary}
-regions, representing invariants that only local capabilties may rely on for
-their safety, and which may be revoked (disabled) in private future worlds.
+valid. Formally, worlds represent the invariants that the global capabilities'
+safety relies on and the reinvocation is only safe in future worlds, where the
+invariants are respected.
+
+However, if we provide the adversary with local capabilities in that first
+invocation, then the situation is a bit different. The adversary has no way to
+store these local capabilities, so if we make sure that there are also no local
+capabilities in the register file for the second invocation (including the
+capability being invoked), then the adversary cannot use them any more.%
+\footnote{We are ignoring write-local capabilities in this discussion. If the
+  adversary does have access to write-local capabilities in the first and second
+  invocation, then the memory they address must entirely be cleared before the
+  second invocation in order for the reinvocation to remain safe.} Therefore, we
+can allow the second invocation to happen in \emph{private} future worlds ($W'
+\futurestr W$), in which global safe capabilities remain safe, but local ones
+don't. This private future world relation is more liberal than the standard,
+public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
+safe). Concretely, worlds may contain \emph{temporary} regions, representing
+invariants that only local capabilties may rely on for their safety, and which
+may be revoked (disabled) in private future worlds.
 
@@
 capabilities (as their safety would be broken in private future worlds), which
-in turn implies that only local capabilities may have store-local permission (a
+in turn implies that only local capabilities may have write-local permission (a
 general sanity requirement when using local capabilities\footnote{In fact, local
   capabilities become useless as soon as the adversary has access to a single
-  global, store-local capability.}).
+  global, write-local capability.})\footnote{For explanation purposes, this
+  discussion ignores certain ways to allow for local capabilities in a permanent
+  region, for example, by not requiring that they are valid or requiring that
+  they are local versions of valid global capabilities etc.}.
 

### 2017-06-15 popl-2018/paper.tex
added possible remark (as a comment) and fixed typo

@@
 aforementioned state transition system where $\Rels$ is a set of pairs
-of relations corresponding to the public and private transition, and
+of relations corresponding to the public and private transitions, and
 $\States$ is a sufficiently large set of states containing at least the
@@
 
-% Notice that the public future region relation is a subset of the
-% private future region relation.
+% not sure whether this should be included:
+% Notice that the public future region relation is a subset of the private future
+% region relation. The public future region relation only allows revoked regions
+% to be reused as temporary region. This means that it is always possible to turn
+% it back into a revoked region in a private future region. This requirement turns
+% out to be important in order to be able to restore the invariants of a caller.
 
@@
 into disjoint parts such that each part is accepted by an active (permanent or
-temporary) region.
+temporary) region. The revoked regions are not taken into account as their memory protocols are no longer in effect.
 \begin{align*}
@@
 \end{align*}
+% Not sure whether to include this remark:
+% It is possible to state nonsensical worlds in which say a global
+% read-write-local capability would be a safe word. There would, however, be no
+% memory that would satisfy such a world which for all intents and purposes would
+% make it discounted by the logical relation.
 

### 2017-06-15 popl-2018/paper.tex
Fix a nuance in the introduction of local caps in the LR after discussion with Lau

@@
 invocation, then the situation is a bit different. The adversary has no way to
-store these local capabilities, so if we make sure that there are also no local
-capabilities in the register file for the second invocation (including the
+store these local capabilities, so if we make sure that there are also no old
+local capabilities in the register file for the second invocation (including the
 capability being invoked), then the adversary cannot use them any more.%

### 2017-06-16 popl-2018/paper.tex
removed comment on limitations of local capabilities as it may not be correct.
First attempt at small argument/local capability discussion

@@
 \end{description}
-
-\textbf{A word on local capabilities} Local capabilities are an essential part
-of the calling convention. The use of local capabilities is, however, not
-limitted to the stack and return pointer.
 
@@
 be global but the callbacks we construct ourselves are allocated on the heap and
-also global. \\\\
+also global.
+
+\emph{Arguments and local capabilities} 
+%Stack and return pointer
+%Local capabilities cannot be "freely used"
+%Safety dictated by LR
+%Need to dynamical check whether we overwrite our local stuff.
+Local capabilities are a central part of the calling convention as they are used
+to construct stack and return pointer. The use of local capabilities for the
+calling convention unfortunately limits the extend of which local capabilties
+can be used for other things. Say we are using the calling convention and
+receive a local capability other than the stack and return pointer, then we need
+to be careful if we want to use it because it may be an alias to the stack
+pointer. That is, if we first push something to the stack and then write to the
+local capability, then we may overwrite our local state. While this sounds
+unsafe, the safety actually depends on what our memory invariants we want to
+establish, so the safety is actually dictated by the logical relation. This may,
+however, not be a satisfactory requirement for say compiler developers as it is
+impossible to figure out what local capabilities are safe to use. A different
+approach would be to do dynamic checks on local capabilities, but this may be to
+costly to do in practice.
+
+%Pass capability on
+%Stack example
+%Pass argument by reference
+%Sofisticated deep safety analysis
+We also need to be careful when we receive a capability from an adversary that
+we want to pass on to a different adversary. It turns out that the logical
+relation again dictates when this is safe. Namely, the logical relation says
+that we can only pass on safe arguments. For instance, when we receive a stack
+pointer from an adversary, then we may at some point want to pass on part of
+this stack pointer to say a callback. In order to do so, we need to make sure
+the stack pointer is safe which means that it must only point to safe
+values. When received from an adversary, we have to consider the contents of the
+stack unsafe, so before we pass it on we clear which corresponds to filling it
+with safe values as integers are always safe to pass on. In general, if we want
+to pass on a local capability we received from an adversary, we have to make
+sure it is safe, for instance by clearing everything it points to (if possible,
+the capability might be a readonly capability). Clearing everthing is, however,
+not always a desirable approach as we may have an array or data structure with
+important values we want to pass on by reference in which case it is
+unacceptable to just clear the entire thing. In this case, it might be possible
+to make a deep safety analysis to make sure that the capability points to
+nothing dangerous. What we are left with is a too conservative approach or an
+non-obvious expensive dynamic safety analysis.
+
+% Can local capabilities be used for other things?/Local capabilities only for stack?
+%Not unheard of - other machine also have things specific for language features.
+The use of local capabilities for other things than stack and return pointer
+turns out to be somewhat limitted. It is, however, not unusual that machines
+have builtin special constructs that are used for just one language feature. A
+machine could for instance have call and return instructions with the sole
+prupose of handling everything related to calls.
+\\\\
+% Old version of the above:
+% Carefully constructing stack and return
+% pointer plays a big part in the calling convention, but it is completely
+% undermined if an unsafe argument, say the full stack pointer that include the
+% caller's local stack frame, is passed as an argument of the call. The logical
+% relation dictates when it is safe to pass an argument to an adversary, namely it
+% must be a safe value. In order to illustrate what this means for local
+% capabilities, consider the stack capability. When the stack capability is
+% received from an adversary, then the contents of the stack is unknown and must
+% thus be considered unsafe, so before the stack capability can be passed on, we
+% must make sure that it only gives access to safe values. In order to meet this
+% requirement, we take a conservative approcah and simply clear everything on the
+% stack. Integers are always safe, so the stack capability can now safely be
+% passed on. While this approach works for the stack, it may in practice be a bit
+% too naive for other local capabilities received form an adversary. Say we
+% receive a capability from an adversary that happens to be the first part of the
+% stack. We want to pass this capability on, so we naively clear the contents it
+% gives access to which in the worst case means we overwrite our local data or the
+% activation record. While this is strictly speaking not unsafe (integers are
+% always) safe, it is not desirable in a real system. We would in principle also
+% like to be able to pass on arguments by reference, e.g a local capability for
+% some array of values. In this case, it would not be satisfactory to just clear the
+% entire thing. However, the capability still needs to be safe, so there would
+% need to be some kind of deep dynamic check that ensures that everything the
+% capability gives access to is indeed safe. In a real setting, this may again be
+% undesirable.
+% This also points to a more
+% general problem: When can we use local capabilities we have received from an
+% adversary. We at least need to do some kind of dynamic range check to make sure
+% that we do not inadverdently overwrite our local stack frame.
 

### 2017-06-16 popl-2018/paper.tex
Added a couple of remarks I was on the fence about

@@
 
-% not sure whether this should be included:
-% Notice that the public future region relation is a subset of the private future
-% region relation. The public future region relation only allows revoked regions
-% to be reused as temporary region. This means that it is always possible to turn
-% it back into a revoked region in a private future region. This requirement turns
-% out to be important in order to be able to restore the invariants of a caller.
+Notice that the public future region relation is a subset of the private future
+region relation. The public future region relation only allows revoked regions
+to be reused as temporary region. This means that it is always possible to turn
+it back into a revoked region in a private future region. This requirement turns
+out to be important in order to be able to restore the invariants of a caller.
 
@@
 \end{align*}
-% Not sure whether to include this remark:
-% It is possible to state nonsensical worlds in which say a global
-% read-write-local capability would be a safe word. There would, however, be no
-% memory that would satisfy such a world which for all intents and purposes would
-% make it discounted by the logical relation.
+It is possible to state nonsensical worlds in which say a global
+read-write-local capability would be a safe word. There would, however, be no
+memory that would satisfy such a world which for all intents and purposes would
+make it discounted by the logical relation.
 

### 2017-06-16 popl-2018/paper.tex
Added a bit to beginning of LR subsection.

@@
 %   respect them itself too.} 
-The logical relation now defines semantically when values, terms, and
-configurations can safely be used or executed by the adversary, i.e. without the
-risk that they will be able to use it to break protocols in the world. The
-definition of the logical relation is found in \figurename{}s~\ref{fig:logrel}
-and~\ref{fig:perm-conds} and we provide some explanations in the following
-paragraphs. We omit some definitions and explain them only verbally, but precise
-definitions can be found in the technical appendix~\citep{technical_appendix}.
+The logical relation defines semantically when values, program counters, and
+configurations are safe. In Section~\ref{sec:formalizing-guarantees}, we
+introduced what is meant by safe values. Similarly for program counters and
+configurations, they are safe with respect to a world when they stick to the
+memory protocols described by the world.
+
+The definition of the logical relation is found in
+\figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
+explanations in the following paragraphs. We omit some definitions and explain
+them only verbally, but precise definitions can be found in the technical
+appendix~\citep{technical_appendix}.
 

### 2017-06-16 popl-2018/paper.tex
finished first read through of LR section - still contemplating whether the last part about execute condition should be updated to talk about an adversary

@@
 % Uncomment me to disable comments
-\renewcommand\lau[1]{}
-\renewcommand\dominique[1]{}
-\renewcommand\lars[1]{}
+% \renewcommand\lau[1]{}
+% \renewcommand\dominique[1]{}
+% \renewcommand\lars[1]{}
 
@@
 
-The \emph{observation relation} $\observations$ defines configurations that we
-consider safe. Roughly speaking, a configuration is safe if whenever it halts,
-there is a private future world that the resulting memory satisfies. Notice that
-failing is considered safe behavior. In fact, the machine often resorts to
-failing when an unauthorized access is attempted, like loading from a capability
-without read permission. This is similar to \citet{Devriese:2016ObjCap}'s logical
-relation for an untyped language, but unlike typical logical relations
-for typed languages, which require that programs do not fail.
+The \emph{observation relation} $\observations$ defines what configurations we
+consider safe. As previously states, a configuration is safe with respect to a
+world when the execution of said configuration does not break the memory
+protocols of the world. Roughly speaking, this means that when the execution of
+a configuration halts, then there is a private future world that the resulting
+memory satisfies.  Notice that failing is considered safe behavior. In fact, the
+machine often resorts to failing when an unauthorized access is attempted, like
+loading from a capability without read permission. This is similar to
+\citet{Devriese:2016ObjCap}'s logical relation for an untyped language, but
+unlike typical logical relations for typed languages, which require that
+programs do not fail.
 
@@
 program counter if it can be plugged into a safe register file (i.e.\ a register
-file in $\stdrr$) and paired with a memory satisfying the world to produce a
-safe configuration.
+file in $\stdrr$) and paired with a memory satisfying the world to become a
+safe configuration. This means that integers and non-executable capabilities
+(e.g. readonly and enter capabilities) are considered safe as program counters
+because when they are plugged into a register file and paired with a memory, the
+execution will immediately fail, which is safe.
 
 The \emph{value relation} $\stdvr$ defines when words are safe, i.e.\ when they
-cannot be used to break the protocols defined by the world. Non-capability data
-is always safe because it provides no authority. Capabilities give the authority
-to manipulate memory and potentially break memory protocols, so they need to
-satisfy certain conditions to be safe. In \figurename~\ref{fig:perm-conds}, we
-define such a condition for each kind of permission a capability can have (we
-will call these permission-based conditions).
+cannot be used to break the memory protocols defined by the world. We make the
+value relation as liberal as possible by considering what is the most we can
+allow an adversary to use a capability for without breaking the memory
+protocols. Non-capability data is always safe because it provides no
+authority. Capabilities give the authority to manipulate memory and potentially
+break memory protocols, so they need to satisfy certain conditions to be
+safe. In \figurename~\ref{fig:perm-conds}, we define such a condition for each
+kind of permission a capability can have (we will call these permission-based
+conditions).
 
 For capabilities with read permission, the $\readCond{}$ ensures that it can
-only be used to read safe words, by requiring that the region governing the
-relevant part of memory, guarantees that it contains only words in the value
-relation. Generally speaking, global capabilities should only depend on
-permanent regions. Therefore we use the function $\var{localityReg}(\gl,W)$,
-which projects out all active regions when $\gl$ (the locality) is local, but
-only the permanent regions when $\gl$ is global. The read condition is defined
-in terms of a standard region $\iota^\pwl_{\start,\addrend}$ which requires all
-the words in the range $[\start,\addrend]$ to be safe.
-
-For a capability with write permission, $\writeCond{}$ must be
-satisfied for the capability's range of authority. Such a capability
-may be used to write any word, so the region governing the relevant
-memory must (at least) allow any safe word to be stored. More
-precisely, for write-local capabilities it needs to allow arbitrary
-safe values to be stored, while for write capabilities, storing values
-that are both non-local and safe is sufficient.  This is done in the
-value relation by the regions the write condition is applied with. For
-write-local capabilities $\iota^\pwl$ is used, and for write
-capabilities we use another standard region
-$\iota^\nwl_{\start,\addrend}$ which requires that the words in
+only be used to read safe words by putting an upper bound on what can be
+required by the region governing the memory the read capability gives access
+to. An adversary can read anything the capability governs, so all the words they
+can possibly read must be safe - that is they must be in the value relation. The
+read condition also depends on the locality of the capability because, generally
+speaking, global capabilities should only depend on permanent
+regions. Concretely to make sure global capabilities only depend on permanent
+regions, we use the function $\var{localityReg}(\gl,W)$, which projects out all
+active regions when $\gl$ (the locality) is local, but only the permanent
+regions when $\gl$ is global. The upper bound on the region is enforced in terms
+of a standard region $\iota^\pwl_{\start,\addrend}$ which requires all the words
+in the range $[\start,\addrend]$ to be safe.  The region in the world that
+actually governs the part of memory the capability gives access to must as most
+be as restrictive as the standard region.  \lau{Here I would have liked to point
+  back to the bit about step-indexing in the intuition section, but
+  unfortunately the ``step down'' happens in the standard region which we have
+  not included.}
+
+For a capability with write permission, $\writeCond{}$ must be satisfied for the
+capability's range of authority. An adversary can use such a capability to write
+any word they can get a hold of, and we can safely assume that they can get a
+hold of any safe word, so the region governing the relevant memory must allow
+any safe word to reside there. In the name of making the logical relation as
+liberal as possible, we make this a lower bound of what the region may allow.
+\lau{Maybe add: Conceptually, the lower bound also conveys the amount of faith
+  one can put in the contents of a piece of memory that an adversary has been
+  given write access to.}  For write capabilities, we also have to take into
+account that we have two flavours of write permissions namely write and
+write-local. In the case of write-local capabilities, the region needs to at
+least require any safe word to be written, but in the case of write
+capabilities, the capability cannot be used to write local capabilties, so the
+region needs only to at least require safe non-local values.  In the write
+condition, this is handled by parameterising it with a region. For the
+write-local capabilities the write condition is applied with the standard region
+$\iota^\pwl$ and for the write capabilities we use a different standard region
+namely $\iota^\nwl_{\start,\addrend}$ which requires that the words in
 $[\start,\addrend]$ are non-local and safe. Like before, we use
-$\var{localityReg}$ to pick an appropriate region for the capability's
-locality. The condition for the region to be address-stratified is a
-technical requirement, defined in the technical appendix.
-An interesting property of the read and write condition is that they
-prohibit global write-local capabilities which, as discussed in
-Section~\ref{sec:stack-and-return-pointer}, is necessary for the safe
-use of local capabilities.
+$\var{localityReg}$ to pick an appropriate region based on the capability's
+locality. Finally the region is required to be address-stratified. Intuitively,
+this means that if a region accepts two memory segments, then it must also
+accept every memory segment ``inbetween'' that is every memory segment where
+each address contains a value from one of the two memory segments. An
+interesting property of the read and write condition is that they prohibit
+global write-local capabilities which, as discussed in
+Section~\ref{sec:stack-and-return-pointer}, is necessary for the safe use of
+local capabilities.
 

### 2017-06-18 popl-2018/paper.tex
intro pass

--- 2017-06-16 popl-2018/paper.tex
@@
 \begin{itemize}
-\item Formalize a simple but representative capability machine featuring local
+\item We formalize a simple but representative capability machine featuring local
   capabilities, and its operational semantics
-  (Section~\ref{sec:capab-mach-with})
-\item Define a novel calling convention enforcing control-flow correctness and
+  (Section~\ref{sec:capab-mach-with}).
+\item We define a novel calling convention enforcing control-flow correctness and
   encapsulation of stack frames (Section~\ref{sec:stack-and-return-pointer}). It
@@
   support: an efficient instruction for clearing a range of memory.
-\item A novel Kripke logical relation for reasoning about programs on the
-  capability machine. The fundamental theorem of logical relations here gives
-  the most general and powerful formulation of the guarantees offered by a
+\item We present a novel Kripke logical relation for reasoning about programs on the
+  capability machine. For our logical relation we prove an analogue of
+  the standard fundamental theorem of logical relations --- our theorem
+  gives  the most general and powerful formulation of the formal guarantees offered by a
   capability machine (a form of capability
@@
   our calling convention or a specific way of using the system's capabilities.
-\item Introduce two novel technical ideas in the unary, step-indexed Kripke
+\item We introduce two novel technical ideas in the unary, step-indexed Kripke
   logical relation used to formulate the above theorem. First, we model the
-  naturally continuation-passing-style of assembly code (with return pointers as
-  continuations) using a \emph{single} orthogonal closure, rather than the
-  existing biorthogonal closure. Secondly, we express the special nature of
+  natural continuation-passing-style of assembly code (with return pointers as
+  continuations) using a \emph{single} orthogonal closure, rather than
+  the earlier used biorthogonal closure. Second, we express the special nature of
   local capabilities using a variant of Dreyer et. al.'s public and private
-  future worlds~\citep{Dreyer:jfp12}, used to express well-bracketedness
-  guarantees in a lambda calculus. The logical relation and capability safety
-  theorem are presented in Section~\ref{sec:logical-relation}.
-\item Demonstrate our results by applying them to challenging examples,
+  future worlds~\citep{Dreyer:jfp12}, which Dreyer et. al.\ used to express well-bracketedness
+  guarantees in a lambda calculus. The logical relation and the
+  fundamental theorem expressing capability safety
+  are presented in Section~\ref{sec:logical-relation}.
+\item We demonstrate our results by applying them to challenging examples,
   specifically constructed to demonstrate local-state encapsulation and
@@
 \powerset{\MemSegments}$ (which defines which memory segments are safe in the
-current state of the world) is only required to be monotonous w.r.t.\ public
+current state of the world) is only required to be monotone w.r.t.\ public
 future worlds (i.e.\ safe memory segments remain safe in public future worlds),
-but for permanent regions, it must be monotonous w.r.t.\ private future worlds.
+but for permanent regions, it must be monotone w.r.t.\ private future worlds.
 As a consequence, the memory for a permanent region may not contain local

### 2017-06-18 popl-2018/paper.tex
intro

@@
   capability machine. For our logical relation we prove an analogue of
-  the standard fundamental theorem of logical relations --- our theorem
-  gives  the most general and powerful formulation of the formal guarantees offered by a
+  the standard fundamental theorem of logical relations --- to the
+  best of our knowledge, our theorem
+  is % by far
+  the most general and powerful formulation of the formal guarantees offered by a
   capability machine (a form of capability

### 2017-06-19 popl-2018/paper.tex

--- 2017-06-16 popl-2018/paper.tex
@@
   support: an efficient instruction for clearing a range of memory.
-\item A novel Kripke logical relation for reasoning about programs on the
-  capability machine. The fundamental theorem of logical relations here gives
-  the most general and powerful formulation of the guarantees offered by a
+\item We present a novel Kripke logical relation for reasoning about programs on the
+  capability machine. For our logical relation we prove an analogue of
+  the standard fundamental theorem of logical relations --- to the
+  best of our knowledge, our theorem
+  is % by far
+  the most general and powerful formulation of the formal guarantees offered by a
   capability machine (a form of capability

### 2017-06-19 popl-2018/paper.tex
Wrote a bit more in FTLR section: Link back to safety, refer forward to examples.

@@
 preserve the invariants in $W$ under any interaction), then an execution will
-necessarily also preserve the invariants (irrespective of the instructions being
+necessarily also preserve the memory invariants (irrespective of the instructions being
 executed).
@@
 our logical relation: a strong theorem that formalizes the guarantees offered by
-the capability machine. Essentially, it says a capability that only grants
-safe authority is capability safe as a program counter.
+the capability machine. Essentially, it says a capability that only grants safe
+authority is capability safe as a program counter (where safety, as in the
+previous sections, means respect of memory protocols).
 \begin{theorem}[Fundamental Theorem]
@@
 \end{theorem}
+The permission based conditions of Theorem~\ref{thm:ftlr} make sure that the
+capability only provides safe authority in which case the capability must be in
+the $\stder$ relation, i.e. it can safely be used as a program counter in an
+otherwise safe register-file.
+
 The Fundamental Theorem can be understood as a general expression of the
@@
 \emph{any} instruction is fine and will not be able to go beyond the authority
-of the values it has access to.
+of the values it has access to. We demonstrate this in
+Section~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
+capabilities that points to arbitrary code. The relation between
+Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow
+correctness, will also be shown by example in Section~\ref{sec:examples} as the
+examples depend on these properties for correctness.
 

### 2017-06-20 local-capability-machine/local-capability-machine.tex
updated TR with macro implementations (still missing call)

@@
 \end{align*}
+\lau{Our instruction set could use a less than operator.}
 \lau{Dominique, apparently there was a reason I had not added the new condition in store to $\writeAllowed{}$.
@@
               \begin{description}
-              \item[\texttt{fetch} $r$ $f$] load the entry of the linking table corresponding to $f$ to register $r$.
+              \item[\texttt{fetch} $r$ $f$] load the entry of the linking table corresponding to $f$ to register $r$.\\
+One possible malloc implementation (r\_t1 and r\_t2 are registers in RegName\_t).
+\begin{lstlisting}
+move r pc
+getb r_t1 r
+geta r_t2 r
+minus r_t1 r_t1 r_t2 // Offset to first address, i.e., linking table (b-a)
+lea r r_t1
+load r r
+lea r ... // ... replaced with offset to f in the linking table
+move r_t1 0
+move r_t2 0
+load r r // f capability loaded to register r
+\end{lstlisting}
               \item[\texttt{call} $r(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline
@@
               \item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register $r$. 
-% Suggestion for malloc impl (r_t1 and r_t2 are registers in RegName_t), r_1 should not contain important stuff
-% move r_1 n
-% move r pc
-% getb r_t1 r
-% geta r_t2 r
-% minus r_t1 r_t2 r_t1
-% lea r r_t1
-% move r_0 r_t1
-% lea r_0 3 // 3 is offset to just after jmp
-% restrict r_0 encodePerm(e)
-% jmp r
-% move r r_1
-% move r_1 0
-% move r_0 r_t1
-% move r_t1 r_0
-
+One possible malloc implementation (r\_t1 is a register in RegName\_t) and r\_1 is the register from the malloc specification.
+\begin{lstlisting}
+fetch r malloc
+move r_1 n 
+// save return pointer
+move r_t1 r_0 
+// setup new return pointer
+move r_0 pc
+lea r_0 4 // 4 is the offset to just after jmp r
+restrict r_0 encodePerm(e)
+jmp r
+move r r_1
+move r_0 r_t1 // restore return pointer
+move r_1 0
+move r_t1 0
+\end{lstlisting}
               \item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison).
 If they are equal, then execution continues.
-If they are unequal, then the assertion flag named $\var{flag}$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
+If they are unequal, then the assertion flag named $\var{flag}$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).\\
+There are four different asserts based on whether $r_1$ and $r_2$ are registers or numbers. If $r_1$ and $r_2$ are registers:
+\begin{lstlisting}
+          // setup pointer to fail.
+          move r_t3 pc
+          lea r_t3 ... // ... is the offset to fail
+          // make sure both registers contain either capability or integer
+          isptr r_t1 r_1
+          isptr r_t2 r_2
+          minus r_t1 r_t1 r_t2
+          jnz   r_t3 r_t1
+          // set up capability for cap case:
+          move r_t4 pc
+          lea r_t4 ... // ... is the offset to caps
+          jnz r_t4 r_t2 // jump to caps if r_t2 contains a capability
+          // the two registers contain an integer
+          minus r_t1 r_1 r_2
+          jnz r_t3 r_t1
+          // the two integers in the registers are equal
+          move r_t4 pc
+          lea r_t4 ... // .. offset to success
+caps:
+          geta r_t1 r_1
+          geta r_t2 r_2
+          minus r_t1 r_t1 r_t2
+          jnz   r_t3 r_t1          
+          getb r_t1 r_1
+          getb r_t2 r_2
+          minus r_t1 r_t1 r_t2
+          jnz   r_t3 r_t1          
+          gete r_t1 r_1
+          gete r_t2 r_2
+          minus r_t1 r_t1 r_t2
+          jnz   r_t3 r_t1          
+          getp r_t1 r_1
+          getp r_t2 r_2
+          minus r_t1 r_t1 r_t2
+          jnz   r_t3 r_t1          
+          getl r_t1 r_1
+          getl r_t2 r_2
+          minus r_t1 r_t1 r_t2
+          jnz   r_t3 r_t1          
+          // the two capabilities in the registers are equal
+          move r_t4 pc
+          lea r_t4 ... // .. offset to success
+fail:
+          // get the flag capability
+          move r_t3 pc
+          getb r_t1 pc
+          geta r_t2 pc
+          minus r_t1 r_t1 r_t2
+          lea r_t3 r_t1
+          lea r_t3 1 // the flag table capability is at the second address of cap.
+          load r_t1 r_t3
+          lea r_t1 ... // ... is the offset of flag in the table
+          store r_t1 1
+          halt
+success:                   
+          // clean up
+          move r_t1 0
+          move r_t2 0
+          move r_t3 0
+          move r_t4 0
+\end{lstlisting}
+If $r_1$ is a register, but $r_2$ is a constant:
+\begin{lstlisting}
+          // setup pointer to fail.
+          move r_t3 pc
+          lea r_t3 ... // ... is the offset to fail
+          // make sure both registers contain either capability or integer
+          isptr r_t1 r_1
+          jnz   r_t3 r_t1
+          minus r_t1 r_1 r_2
+          jnz r_t3 r_t1
+          // the two integers in the registers are equal
+          move r_t3 pc
+          lea r_t3 ... // .. offset to success
+fail:
+          // get the flag capability
+          move r_t3 pc
+          getb r_t1 pc
+          geta r_t2 pc
+          minus r_t1 r_t1 r_t2
+          lea r_t3 r_t1
+          lea r_t3 1 // the flag table capability is at the second address of cap.
+          load r_t1 r_t3
+          lea r_t1 ... // ... is the offset of flag in the table
+          store r_t1 1
+          halt
+success:                   
+          // clean up
+          move r_t1 0
+          move r_t3 0
+\end{lstlisting}
+The case where $r_1$ is a constant and $r_2$ is a register is omitted.
+The case where both are constant is also omitted - if the constants are the same, then the macro is nothing.
+If they are different, then it corresponds to the failed part of both of the above implementations.
               \item[\texttt{mclear $r$}] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction.
-In a real system it would probably be implemented as a vector operation which allows modification of continuous segments of memory rather fast.}
-              \item[\texttt{rclear $\bar{r}$}] Moves 0 to all the registers in the list $\bar{r}$.
+In a real system it would probably be implemented as a vector operation which allows modification of continuous segments of memory rather fast.} \\
+Possible implementation:
+\begin{lstlisting}
+        move r_t r
+        getb r_t1 r_t
+        geta r_t2 r_t
+        minus r_t2 r_t1 r_t2
+        lea r_t r_t2
+        gete r_t2
+        minus r_t1 r_t2 r_t1
+        plus r_t1 r_t1 1
+        move r_t2 pc
+        lea r_t2 ... // ... is the offset to end
+        move r_t3 pc
+        lea r_t3 ... // ... is the offset to iter
+iter:
+        jnz r_t2 r_t1
+        store r_t 0
+        lea r_t 1
+        plus r_t1 r_t1 1
+        jmp r_t3
+end:
+        move r_t 0
+        move r_t1 0
+        move r_t2 0
+        move r_t3 0
+\end{lstlisting}
+              \item[\texttt{rclear $\bar{r}$}] Moves 0 to all the registers in the list $\bar{r}$.\\
+Possible implementation: Say $\bar{r} = r_1,\dots, r_n$
+\begin{lstlisting}
+move r_1 0
+move r_2 0
+// ...
+move r_n 0
+\end{lstlisting}
+
 \end{description}
@@
 \begin{description}
-              \item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the address of the stack capability by one and storing the word through the stack capability.
+\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by incrementing the address of the stack capability by one and storing the word through the stack capability.\\
+Possible implementation:
+\begin{lstlisting}
+  lea r_stk 1
+  store r_stk r
+\end{lstlisting}
+
 \item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, and decrementing the address of the stack capability.
+\begin{lstlisting}
+  load r r_stk
+  minus r_t1 0 1
+  lea r_stk r_t1
+\end{lstlisting}
+
 \item[\texttt{scall} $r(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] \forcenewline$\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers.
@@
     \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
-                  \end{itemize}\
+    \end{itemize}
   \end{description}
 \end{description}
+Possible implementation, say $\bar{r}_{\var{args}} = r_{\var{args},1},\dots, r_{\var{args},m}$ and $\bar{r}_{\var{priv}} = r_{\var{priv},1},\dots, r_{\var{priv},n}$:
+\begin{lstlisting}
+// push private state
+  push r_priv,1
+  ...
+  push r_priv,n
+// push activation code
+  push encode(i_1)
+  ...
+  push encode(i_4)
+// push old pc
+  move r_t1 pc
+  lea r_t1 ... // ... is the offset to after
+  push r_t1
+// push stack pointer
+  push r_stk
+// set up protected return pointer
+  move r_0 r_stk
+  lea r_0 -5 // -5 is the offset to the first instruction of the activation code
+  restrict r_0 encodePermPair((Local,e))
+// restrict stack capability
+  geta r_t1 r_stk
+  plus r_t1 r_t1 1
+  getb r_t2 r_stk
+  subseg r_stk r_t1 r_t2
+// clear unused part of the stack
+  mclear r_stk
+// clear non-argument registers
+  rclear R // where R = RegisterName - {pc,r_stk,r_0,r,r_args}
+  jmp r
+after:
+// pop the restore code
+  pop r_t1
+  pop r_t1
+  pop r_t1
+  pop r_t1
+// pop the private state into approriate registers
+  pop r_priv,1
+  ...
+  pop r_priv,n
+
+\end{lstlisting}
+where the restore code is as follows:
+\begin{lstlisting}
+ i_1 = move r_t1 pc
+ i_2 = lea r_t1 5 // 5 is the offset to the address where the old stack pointer is located
+ i_3 = load r_stk r_t1
+ i_4 = pop pc
+\end{lstlisting}
 Note:
@@
   \end{description}
-%Possible implementation of crtcls \overline{(x,r_v)} r_code where |\overline{(x,r_v)}| = n (i_1,...,i_6 under)
-% malloc r_t1 n
-% store r_t1 r_v1
-% lea r_t1 1
-% store r_t1 r_v2
-% lea r_t1 1
-% ...
-% lea r_t1 1
-% store r_t1 r_vn
-% lea r_t1 -n
-% restrict r_t1 encodePerm(rw)
-% malloc r_1 8 //length of activation record
-% store r_1 r_code
-% lea r_1 1
-% store r_1 r_t1
-% move r_t1 0
-% lea r_1 1
-% store r_1 encode(i_1)
-% lea r_1 1
-% store r_1 encode(i_2)
-% lea r_1 1
-% ...
-% lea r_1 1
-% store r_1 encode(i_6)
-% lea r_1 -5 //offset to first instruction
-% restrict r_1 encodePerm(e)
-
-%Activation code (i_1,...,i_6)
-% i_1 = move r_t1 pc
-% i_2 = lea r_t1 -2
-% i_3 = load r_env r_t1
-% i_4 = lea r_t1 1
-% i_5 = load r_t1 r_t1
-% i_6 = jmp r_t1
+Possible implementation of \texttt{crtcls $\overline{(x,r_v)}$ $r_{\var{code}}$} where $|\overline{(x,r_v)}| = n$ ($i_1$,...,$i_6$, i.e. the activation code, is defined later):
+\begin{lstlisting}
+malloc r_t1 n
+store r_t1 r_v1
+lea r_t1 1
+store r_t1 r_v2
+lea r_t1 1
+...
+lea r_t1 1
+store r_t1 r_vn
+lea r_t1 -n
+restrict r_t1 encodePermPair((Global,rw))
+malloc r_1 8 //length of activation record
+store r_1 r_code // code capability
+lea r_1 1
+store r_1 r_t1 // environment capability
+move r_t1 0
+lea r_1 1
+store r_1 encode(i_1)
+lea r_1 1
+store r_1 encode(i_2)
+lea r_1 1
+...
+lea r_1 1
+store r_1 encode(i_6)
+lea r_1 -5 //offset to first instruction
+restrict r_1 encodePerm(e)
+\end{lstlisting}
+Activation code ($i_1$,...,$i_6$):
+\begin{lstlisting}
+i_1 = move r_t1 pc
+i_2 = lea r_t1 -2
+i_3 = load r_env r_t1
+i_4 = lea r_t1 1
+i_5 = load r_t1 r_t1
+i_6 = jmp r_t1
+\end{lstlisting}
 
@@
 Loads the word at the index associated with $x$ in the environment list.
-Loads from this capability into $r$.
+Loads from this capability into $r$.\\
+Possible implementation:
+\begin{lstlisting}
+  move r_t1 r_env
+  lea r_t1 ... // ... corresponds to offset of x in environment
+  load r r_t1
+  move r_t1 0
+\end{lstlisting}
 \item[\texttt{store $x$ $r$}] Assumes environment capability available in register $r_{\var{env}}$.
@@
 Stores the contents of register $r$ through this capability.
-              \item[\texttt{reqglob $r$}] Tests if register $r$ contains a $\glob$ capability. If not fail, otherwise continue execution.
-              \item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with permission $\decodePerm{n}$. If not fail, otherwise continue execution.
+\begin{lstlisting}
+  move r_t1 r_env
+  lea r_t1 ... // ... corresponds to offset of x in environment
+  store r_t1 r
+  move r_t1 0
+\end{lstlisting}
+\item[\texttt{reqglob $r$}] Tests if register $r$ contains a $\glob$ capability. If not fail, otherwise continue execution.\\
+Possible implementation:
+\begin{lstlisting}
+  getl r_t1 r
+  minus r_t1 r_t1 encodeLoc(Global)
+  move r_t2 pc
+  lea r_t2 4 // 4 is the offset to just after fail
+  jnz r_t1 r_t2
+  fail
+  move r_t1 0
+  move r_t2 0
+\end{lstlisting}
+\item[\texttt{reqperm $r$ $n$}] Tests if register $r$ contains a capability with permission $\decodePerm{n}$. If not fail, otherwise continue execution.\\
+Possible implementation:
+\begin{lstlisting}
+  getp r_t1 r
+  minus r_t1 r_t1 n
+  move r_t2 pc
+  lea r_t2 4 // 4 is the offset to just after fail
+  jnz r_t1 r_t2
+  fail
+  move r_t1 0
+  move r_t2 0
+\end{lstlisting}
+
 \item[\texttt{prepstack} $r$] Tests if register $r$ contains a capability with permission $\rwlx$.
-If not fail, otherwise assume $r$ points to $((\rwlx,\gl),\start,\addrend,\addr)$ adjust it to $((\rwlx,\gl),\start,\addrend,\start - 1)$.
-                \lau{What if the stack starts at address 0? Changing the stack convention to always point at the first free cell will get rid of this problem (because our memory us uncapped.)}
-              \item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 
+If not fail, otherwise assume $r$ points to $((\rwlx,\gl),\start,\addrend,\addr)$ adjust it to $((\rwlx,\gl),\start,\addrend,\start - 1)$.\\
+Possible implementation
+\begin{lstlisting}
+  reqperm r encodePerm(rwlx)
+  getb r_t1 r
+  geta r_t2 r
+  minus r_t1 r_t1 r_t2
+  lea r r_t1
+  minus r_t1 0 1
+  lea r r_t1
+  move r_t1 0
+  move r_t2 0
+\end{lstlisting}
+  \lau{What if the stack starts at address 0?
+Changing the stack convention to always point at the first free cell will get rid of this problem (because our memory us uncapped.)}\lau{Seeing as we did not change this: It does not give a problem as such because this would cause the machine to fail which is considered acceptable.
+It would also be possible to make a more ``sofisticated'' implementation that special cases on this and simply throws the first address away.}
 \end{description}
@@
 % \subsubsection{Simple use of local capabilities}
-%               The following programs only use local capabilities when they create protected return pointers.
-The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
+% The following programs only use local capabilities when they create protected return pointers. The strength of this is, however, not shown as the untrusted code only gets one chance to execute.
 
@@
     \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
-    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{c_\adv},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r_\adv \mapsto c_\adv]$ 
+    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{c_\adv},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk', r_\adv \mapsto c_\adv]$ 
     \item $\npair[n-k-1]{c_{\mathit{ret}}} \in \stdvr(W')$
@@
     \begin{enumproof}
-      \item $\npair[n-k-1]{c_\adv} \in \stdvr{W}$ \\
+      \item $\npair[n-k-1]{c_\adv} \in \stdvr(W)$ \\
         To this end let $n' < n-k-1$ and $W'' \futurestr W'$ be given and show
@@
 \begin{lstlisting}
-g2:  move $\pcreg$ $r_3$
+g2:  move $r_3$ $\pcreg$
      lea $r_3$ $\dots$
@@
   &\iota^\pwl : \cal{P} \fun \Regions\\
-  &\iota^\pwl\;A \defeq (\temp,1,=,=,H^\pwl \; S) \\
+  &\iota^\pwl\;A \defeq (\temp,1,=,=,H^\pwl \; A) \\
   \\

### 2017-06-21 local-capability-machine/local-capability-machine.tex
Added call macro - section could still use a bit of polishing.

@@
                 \end{description} 
+Possible implementation. We will use $\texttt{malloc $r$ $n$}$ and \texttt{rclear $\bar{r}$} (defined below). Assume $\bar{r_{\var{priv}}} = r_{\var{priv},1}, \dots, r_{\var{priv},n}$
+\begin{lstlisting}
+  malloc r_t ... // ... is the size of activation record
+// store private state in activation record
+  store r_t r_priv,1
+  lea r_t 1
+  store r_t r_priv,2
+  lea r_t 1
+  ...
+  lea r_t 1
+  store r_t r_priv,n
+  lea r_t 1
+// store old pc
+  move r_t1 pc
+  lea r_t1 ... // ... is the offset to return address
+  store r_t r_t1
+  lea r_t1 1
+// store activation record
+  store r_t encode(i_1)
+  lea r_t1 1
+  ...
+  lea r_t1 1  
+  store r_t encode(i_m)
+  lea r_t1 k //  k is m-1, i.e. the offset to the first instruction of the activation code.
+  restrict r_t1 encodePermPair((Local,e))
+  move r_0 r_t1
+  rclear R // R = RegisterName - {r,pc,r_0,r_args}
+  jmp r
+\end{lstlisting}
+Activation record. The instructions correspond to $i_1,\dots,i_m$ in the above.
+\begin{lstlisting}
+  move r_t pc
+  getb r_t1 r_t
+  geta r_t2 r_t
+  minus r_t1 r_t1 r_t2
+// load private state
+  lea r_t r_t1
+  load r_priv,1 r_t
+  lea r_t 1
+  load r_priv,2 r_t
+  lea r_t 1
+  ...
+  lea r_t 1
+  load r_priv,n r_t
+  lea r_t 1
+// load old pc
+  load pc r_t
+\end{lstlisting}
+
               \item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register $r$. 

### 2017-06-21 popl-2018/paper.tex
small part about single stack, please read

@@
 
+\emph{Single stack} A single stack is a good choice for the simple capability
+machine presented here, because it works well with higher-order functions. The
+alternative to a single stack would be to have one stack per component. With
+multiple stacks on this simple capability machine, it is not entirely obvious
+how each component would retrieve their stack pointer. Further with higher-order
+functions, it is not clear which stack the higher-order function should
+use. Should it use the stack of the component that created it or the stack of
+the component in which it is called. If we let it use the stack of the component
+that uses the higher-order function, then what is the purpose of having multiple
+stacks when the stack is not exclusive to each component. If we let it use the
+stack of its creator, then the question is how do you obtain this stack
+pointer. We do not claim that it is impossible to make it work with multiple
+stack in the context of higher-order functions, but it is not obvious how it
+would work.
+
 \noindent \textbf{Reasoning about capability machine programs}\\

### 2017-06-21 popl-2018/paper.tex
Added lt to paper and changed what a word is.

@@
 \newcommand{\minus}[3]{\threeinstr{minus}{#1}{#2}{#3}}
+\newcommand{\lt}[3]{\threeinstr{lt}{#1}{#2}{#3}}
 
@@
     \addr   &\in& \Addrs &\isdef& \nats &\phantom{makesomero}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
-    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words\\
+    w &\in&\Words &\isdef& \ints + \Caps + \{\infty\} &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words\\
     \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
@@
                  \minus{\lv}{\rv}{\rv} \mid 
-                 \lea{\lv}{\rv} \mid \\
-          &   &  \restricttwo{\lv}{\rv} \mid
+                 \lt{\lv}{\rv}{\rv} \mid \\
+          &   &  \lea{\lv}{\rv} \mid 
+                \restricttwo{\lv}{\rv} \mid
                  \subseg{\lv}{\rv}{\rv} \mid  
@@
                  \getp{\lv}{\lv} \mid 
-                 \getl{\lv}{\lv} \mid \getb{\lv}{\lv} \mid
-                 \gete{\lv}{\lv} \mid \\
-          &   &  \geta{\lv}{\lv} \mid 
+                 \getl{\lv}{\lv} \mid 
+                 \getb{\lv}{\lv} \mid \\
+          &   &  \gete{\lv}{\lv} \mid 
+                 \geta{\lv}{\lv} \mid 
                  \fail \mid
@@
 load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and arithmetic addition operators ($\plus{}{}{}$ and
+and $\store{}{}$) and arithmetic addition operators ($\lt{}{}{}$, $\plus{}{}{}$ and
 $\minus{}{}{}$, operating only on numbers). There are three instructions for
@@
 Section~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
-capabilities that points to arbitrary code. The relation between
+capabilities that points to arbitrary instructions. The relation between
 Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow

### 2017-06-21 popl-2018/paper.tex
related, in progress

--- 2017-06-21 popl-2018/paper.tex
@@
 
-\dominique{Should we also discuss the fact that we only have a semantic model
-  but not a logic?}
-
 \section{Related work}
@@
 biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
+Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
+object capability patterns\citep{swasey:2017}. The logic is based on
+Iris~\cite{iris,iris2,iris3}, a state of the art higher-order
+concurrent separation logic and is formalized in Coq, building on the
+Iris Proof Mode for Coq~\cite{ipm}.  
+XXX HERE
 

### 2017-06-21 local-capability-machine/local-capability-machine.tex
first efforts towards adding lt
Updated TR and paper to not have the special lta (as discussed w/ Dominique

@@
 \newcommand{\addrend}{\var{end}}
+\newcommand{\inftyend}{-42}
 \newcommand{\pwlv}{\var{pwl}}
@@
 They are otherwise the same as their non-''permit write local'' counterparts.
+
+As we have $\infty$ as a possible address, but our words cannot express $\infty$.
+We pick $\inftyend$ as a representative for $\infty$ when it is in memory (we could have picked any negative noumber).
+Note that $\inftyend$ is not an address, so for address operations $\inftyend$ only represents $\infty$.
+It is the responsible of the programmer to keep track of what represents addresses (and take necessary precautions).
 
@@
                  \minus{\lv}{\rv}{\rv} \mid 
+                 \lt{\lv}{\rv}{\rv} \mid 
                  \lea{\lv}{\rv} \mid 
@@
                                                           \end{cases}\\
+  \sem{\lt{\refreg{r_1}}{\rv_1}{\rv_2}}(\Phi)               & =
+                                                          \begin{cases}
+                                                            \stdUpdatePc{\updateReg{r_1}{1}} &
+                                                            \arraycolsep=0pt
+                                                            \begin{array}[t]{l}
+                                                              \text{if for $i \in \{1,2\}$}\\
+                                                              \quad\text{$n_i = \rv_i$ or $n_i = \memreg(\rv_i)$}\\
+                                                              \quad\text{and in either case $n_i \in \ints$}\\
+                                                              \quad\text{and $n_1 < n_2$}\\                                                              
+                                                            \end{array}\\
+                                                            \failed & \text{otherwise}
+                                                          \end{cases}
+\\
   \sem{\subseg{\refreg{r}}{\rv_1}{\rv_2}}(\Phi) & = 
@@
                                                 \quad\text{$n_i = \rv_i$ or $n_i = \memreg(\rv_i)$}\\
-                                                \quad\text{and in either case $n_i \in \ints$}\\
-                                                \quad\text{and $\start \leq n_1$ and $n_2 \leq \addrend$}\\
+                                                \quad\text{and in either case $n_1 \in \nats$}\\
+                                                \quad\text{and $\start \leq n_1$}\\
+                                                \quad\text{and $n_2 \leq \addrend$ where $n_2 \in \nats$}\\
+                                                \quad\quad\text{or $n_2=\inftyend$ and $\addrend = \infty$}\\
                                                 \quad\text{and $\perm \neq \entry$}\\
@@
                                               \begin{array}[t]{l}
-                                                \text{if $\memreg(r_2) = ((\_,\_),\_,\addrend,\_)$}
+                                                \text{if $\memreg(r_2) = ((\_,\_),\_,\addrend,\_)$ and $\addrend \neq \infty$}
+                                              \end{array} \\
+                                              \stdUpdatePc{\updateReg{r_1}{\inftyend}} &
+                                              \arraycolsep=0pt
+                                              \begin{array}[t]{l}
+                                                \text{if $\memreg(r_2) = ((\_,\_),\_,\infty,\_)$}
                                               \end{array} \\
@@
 But even if $w$ is just an integer, we still need the capability we write through to have some kind of write permission.}
-\dominique{I see, but maybe writeAllowed could take $w$ and $c$ as an argument, rather than just $\perm$? It is now weird that some conditions for writing are in $\writeAllowed{}$, but not all.}
+\dominique{I see, but maybe writeAllowed could take $w$ and $c$ as an argument, ratheppppr than just $\perm$? It is now weird that some conditions for writing are in $\writeAllowed{}$, but not all.}
 
@@
 \texttt{l:} is a meta level label that can be used to refer to a specific address. When placed on the line of a macro, it refers to the first instruction of this macro.
-
+\clearpage
 \section{Examples}
@@
     \begin{aligned}[t]
-      & \{ \npair{i} \mid i \in \ints \} 
+      & \{ \npair{i} \mid i \in \ints \union \{ \infty \} \} 
       \union \\

### 2017-06-22 popl-2018/paper.tex
Updated TR and paper to not have the special lta (as discussed w/ Dominique
Some minor modifications while I'm making a pass over the paper.

@@
   relies solely on local capabilities and does not require OS support (like a
-  trusted stack or call/return instructions). It also does not use separate
-  per-component stacks but a single, fairly standard, contiguous stack. It
-  supports higher-order cross-component calls (e.g., cross-component function
+  trusted stack or call/return instructions).
+  %It also does not use separate per-component stacks but a single, fairly standard, contiguous stack.
+  %domi: don't mention this here. I think some CSF reviewers misunderstood that we were claiming this as a strength of the approach
+  It supports higher-order cross-component calls (e.g., cross-component function
   pointers) and can be efficient assuming only one additional piece of processor
@@
   capability machine (a form of capability
-  safety~\citet{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
+  safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
   guarantees offered for local capabilities. It is very general and not tied to
@@
   local capabilities using a variant of Dreyer et. al.'s public and private
-  future worlds~\citep{Dreyer:jfp12}, which Dreyer et. al.\ used to express well-bracketedness
+  future worlds~\citep{Dreyer:jfp12}, which they used to express well-bracketedness
   guarantees in a lambda calculus. The logical relation and the
@@
 
-For reasons of space, some details and all proofs have been omitted;
+For reasons of space, some details and all proofs have been omitted; please
 refer to the technical appendix~\citep{technical_appendix} for those.
@@
     \addr   &\in& \Addrs &\isdef& \nats &\phantom{makesomero}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
-    w &\in&\Words &\isdef& \ints + \Caps + \{\infty\} &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words\\
+    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words\\
     \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
@@
 However, there is one exception to the rule above: local capabilities can be
-stored to memory for which we have a capability with write-local authority
+stored to memory through a capability with write-local authority
 (i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
@@
 The basic idea of our approach is simple: we stick to a single, rather standard,
-C stack and register-passed return pointers, much like a standard C calling
-convention. However, to prevent various ways of misusing this basic scheme, we
-put local capabilities to work and take a number of not-always-obvious safety
-measures. In the next paragraphs, we will explain the issues to be considered in
-all the relevant situations: when (1) starting our program, (2) returning to the
-adversary, (3) invoking the adversary, (4) returning from the adversary, (5)
-invoking an adversary callback and (6) having a callback invoked by the
-adversary.
+C stack and register-passed stack and return pointers, much like a standard C
+calling convention. However, to prevent various ways of misusing this basic
+scheme, we put local capabilities to work and take a number of
+not-always-obvious safety measures. In the next paragraphs, we will explain the
+issues to be considered in all the relevant situations: when (1) starting our
+program, (2) returning to the adversary, (3) invoking the adversary, (4)
+returning from the adversary, (5) invoking an adversary callback and (6) having
+a callback invoked by the adversary.
 
@@
 object capability patterns\citep{swasey:2017}. The logic is based on
-Iris~\cite{iris,iris2,iris3}, a state of the art higher-order
+Iris~\citep{iris,iris2,iris3}, a state of the art higher-order
 concurrent separation logic and is formalized in Coq, building on the
-Iris Proof Mode for Coq~\cite{ipm}.  
+Iris Proof Mode for Coq~\citep{ipm}.  
 XXX HERE

### 2017-06-23 popl-2018/paper.tex
related

--- 2017-06-21 popl-2018/paper.tex
@@
 Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
-object capability patterns\citep{swasey:2017}. The logic is based on
-Iris~\cite{iris,iris2,iris3}, a state of the art higher-order
+object capability patterns~\citep{swasey:2017}. The logic is based on
+Iris~\citep{iris,iris2,iris3}, a state of the art higher-order
 concurrent separation logic and is formalized in Coq, building on the
-Iris Proof Mode for Coq~\cite{ipm}.  
-XXX HERE
+Iris Proof Mode for Coq~\citep{ipm}.  OCPL gives a more abstract and
+modular way of proving capability safety for a lambda-calculus (with concurrency)
+compared to the earlier work by \citet{Devriese:2016ObjCap}. 
+In the future we would also like to investigate a new program logic
+for reasoning about capability safety for our capability machine
+model.  We think Iris would also be a natural starting point for such
+an endaveour, since Iris is really a framework, which can be instantiated to
+different programming languages. OCPL was able to leverage existing
+Iris specifications for the high-level language; for our capability
+machine model, however, it would be necessary  to devise new kinds of
+specifications for our low-level programs
+with unstructured control-flow. It is likely that we could get
+inspiration from earlier work on logics for assembly programming
+languages, such as XCAP~\citep{xcap}.
 

### 2017-06-23 popl-2018/paper.tex

--- 2017-06-22 popl-2018/paper.tex
@@
 Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
-object capability patterns\citep{swasey:2017}. The logic is based on
+object capability patterns~\citep{swasey:2017}. The logic is based on
 Iris~\citep{iris,iris2,iris3}, a state of the art higher-order
 concurrent separation logic and is formalized in Coq, building on the
-Iris Proof Mode for Coq~\citep{ipm}.  
-XXX HERE
+Iris Proof Mode for Coq~\citep{ipm}.  OCPL gives a more abstract and
+modular way of proving capability safety for a lambda-calculus (with concurrency)
+compared to the earlier work by \citet{Devriese:2016ObjCap}. 
+In the future we would also like to investigate a new program logic
+for reasoning about capability safety for our capability machine
+model.  We think Iris would also be a natural starting point for such
+an endaveour, since Iris is really a framework, which can be instantiated to
+different programming languages. OCPL was able to leverage existing
+Iris specifications for the high-level language; for our capability
+machine model, however, it would be necessary  to devise new kinds of
+specifications for our low-level programs
+with unstructured control-flow. It is likely that we could get
+inspiration from earlier work on logics for assembly programming
+languages, such as XCAP~\citep{xcap}.
 

### 2017-06-23 popl-2018/paper.tex
Another pass over the paper.

--- 2017-06-22 popl-2018/paper.tex
@@
 %% For double-blind review submission
-\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+\documentclass[anonymous,acmsmall,review]{acmart}\settopmatter{printfolios=true}
 %% For single-blind review submission
@@
 For reasons of space, some details and all proofs have been omitted; please
-refer to the technical appendix~\citep{technical_appendix} for those.
+refer to the technical appendix~\citep{technical_appendix} for
+those.\dominique{upload TA as supplementary material instead?}
 
@@
 with protection against stack smashing attacks (i.e.\ buffer overflows on the
-stack overwriting activation records' contents). Luckily, such protection is
-easy for a reasonable compiler to provide thanks to the capability machine's
-fine-grained memory protection\dominique{explain this point better}.
+stack overwriting activation records' contents). Luckily, the capability
+machine's fine-grained memory protection should make it reasonably easy for a
+compiler to implement such protection, by making sure that only appropriately
+bounded versions of the stack pointer are made available to source language
+code.
 
@@
 have access to upon invocation of the callback. As before, this includes return
-pointers and stack pointers from previous stack frames that they may be
+and stack pointers from previous stack frames that they may be
 trying to illegally use inside the callback.
@@
 
-Our answer is that the adversary should provide it to us, much like we provide
+Our answer is that the adversary should provide it to us, just as we provide
 them with a stack pointer when we invoke their code. However, it is important
 that we do not just accept any capability as a stack pointer but check that it
-is safe to use. Specifically, we check that it is indeed a $\rwlx$ capability. Without
+is safe to use. Specifically, we check that it is indeed an $\rwlx$ capability. Without
 this check, an adversary could potentially get control over our local stack
-frame during a callback by passing us a local $\rwx$ capability to a global data
+frame during a subsequent callback by passing us a local $\rwx$ capability to a global data
 structure instead of a proper stack pointer
@@
 structure to access our local state. To prevent this from happening, we need to
-make sure the stack capability gives $\rwlx$ authority, since the system wide assumption then
+make sure the stack capability carries $\rwlx$ authority, since the system wide assumption then
 tells us that the adversary cannot have global capabilities to our local stack.
@@
 \item[At program start-up] A local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local capabilities.
-\item[Before returning to the adversary] Clear non-return-value registers. Clear the part of the stack we had access to.
+\item[Before returning to the adversary] Clear non-return-value registers. Clear
+  the part of the stack we had access to (not just the part we used).
 \item[Before invoking the adversary] Push activation record to the stack. Create
@@
 
+Note: even though the calling convention is the main application in this paper,
+the logical relation we construct is very general and should be regarded as an
+independent contribution.
+
 \subsection{Formalizing the guarantees of the capability machine}
@@
 as they cannot be used to break invariants. Whether a capability is safe depends
-on the authority that it grants. In the above-described world, a read capability
+on the authority that it carries. In the above-described world, a read capability
 for address $0$ is safe, as it can only be used to read the value $42$, which is
@@
 address $a$? The definition then just says that a read capability for address
-$a$ is safe iff the same read capability for address $a$ is safe. This form of
+$a$ is safe if and only if the same read capability for address $a$ is safe. This form of
 cyclic reasoning is related to similar challenges in languages with recursive
 types or higher-order ML-style references, and a standard solution is
-to use step-indexing: Essentially, the cycle is broken by defining safety up to a
+to use step-indexing \citep{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
 certain number of interaction steps. All words will be considered
@@
 notion of evolvable worlds makes our logical relation into a \emph{Kripke}
-logical relation.
+logical relation \citep{pitts_operational_1998}.
 
@@
 be able to reinvoke the code if we can guarantee that those values are still
-valid. Formally, worlds represent the invariants that the global capabilities'
+valid. Formally, worlds represent the invariants that global capabilities'
 safety relies on and the reinvocation is only safe in future worlds, where the
@@
 a temporary region, the predicate $H \in \Worlds \rightarrow
-\powerset{\MemSegments}$ (which defines which memory segments are safe in the
-current state of the world) is only required to be monotone w.r.t.\ public
+\powerset{\MemSegments}$ (which defines the safe memory segments in the
+current world) is only required to be monotone w.r.t.\ public
 future worlds (i.e.\ safe memory segments remain safe in public future worlds),
@@
   region, for example, by not requiring that they are valid or requiring that
-  they are local versions of valid global capabilities etc.}.
+  they are local versions of valid global capabilities.}.
 
@@
 or global, can depend on these protocols. Protocols imposed by temporary regions
-can be revoked at certain points during execution, on the condition that we
-clear local capabilities when this happens. This means that local capabilities
+can be revoked in private future worlds. Doing this may break the safety of
+local capbalities but not global ones. This means that local capabilities
 can safely depend on the protocols imposed by temporary regions, but global
@@
 % world-dependent. 
-As described in Section~\ref{sec:formalizing-guarantees}, the safety of a world
-depends on the safety of its regions which in turn may depend on the safety of
-the values it contain which depend on the safety of a world. In other words, we
-need recursive worlds. Recursive worlds are a common requirement in Kripke
-models and the following lemma uses the method of
+As described in Section~\ref{sec:formalizing-guarantees}, regions define sets of
+safe memory segments, but this set may itself be world-dependent. In other
+words, our worlds are defined recursively. Recursive worlds are a common
+requirement in Kripke models and the following lemma uses the method of
 \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
@@
 \end{theorem}
-In the above theorem, $\States \times \Rels$ corresponds to the
-aforementioned state transition system where $\Rels$ is a set of pairs
-of relations corresponding to the public and private transitions, and
-$\States$ is a sufficiently large set of states containing at least the
-states we use. The last part of the temporary and permanent regions is
-a state interpretation function that determines what memory segments
-the region permits in each state of its states. The different
-monotonicity requirements in the two interpretation functions reflects
-how permanent regions rely only on permanent protocols whereas
+In the above theorem, $\States \times \Rels$ corresponds to the aforementioned
+state transition system where $\Rels$ contains pairs of relations corresponding
+to the public and private transitions, and $\States$ is an unspecified set that
+we assume to contain at least the states we use in this paper. The last part of
+the temporary and permanent regions is a state interpretation function that
+determines what memory segments the region permits in each state of its states.
+The different monotonicity requirements in the two interpretation functions
+reflects how permanent regions rely only on permanent protocols whereas
 temporary regions can rely on both temporary and permanent protocols.
-$\UPred{\HeapSegments}$ is a step-indexed, downwards closed predicate
-on memory segments.
-
-With the recursive domain equation solved, we use the $\Regions$ from
-Theorem~\ref{thm:world-existence} to define the worlds:
+$\UPred{\HeapSegments}$ is a step-indexed, downwards closed predicate on memory
+segments.
+
+With the recursive domain equation solved, we could take $\Wor$ as our notion of
+worlds, but it is technically more convenient to work with the following
+definition instead:
 \[
@@
 either remain revoked or be replaced by a temporary region. This means that the
-public future world relations allows us to reuse a region that has been revoked
+public future world relations allows us to reinstate a region that has been revoked
 earlier.
@@
 \end{mathpar}
-Here, revocation of temporary regions is allowed. As explained above, this can
-be done by replacing them with a special revoked region, which does not describe
-any memory, but in fact, any other region can also be used to replace both
-revoked and temporary regions. On the other hand, permanent regions cannot be
-masked away. They are only allowed to transition according to the private part
-of the transition system.
+Here, revocation of temporary regions is allowed. In fact, temporary regions can
+be replaced by an arbitrary other region, not just the special $\revoked$.
+Conversely, $\revoked$ regions may also be replaced by any other region. On
+the other hand, permanent regions cannot be masked away. They are only allowed
+to transition according to the private part of the transition system.
 
 Notice that the public future region relation is a subset of the private future
-region relation. The public future region relation only allows revoked regions
-to be reused as temporary region. This means that it is always possible to turn
-it back into a revoked region in a private future region. This requirement turns
-out to be important in order to be able to restore the invariants of a caller.
+region relation.
+%domi: the below seems hard to understand at this moment.
+% The public future region relation only allows revoked regions
+% to be reused as temporary region. This means that it is always possible to turn
+% it back into a revoked region in a private future region. This requirement turns
+% out to be important in order to be able to restore the invariants of a caller.
 
@@
 into disjoint parts such that each part is accepted by an active (permanent or
-temporary) region. The revoked regions are not taken into account as their memory protocols are no longer in effect.
+temporary) region.  Revoked regions are not taken into account as their memory protocols are no longer in effect.
 \begin{align*}
@@
 \end{align*}
-It is possible to state nonsensical worlds in which say a global
-read-write-local capability would be a safe word. There would, however, be no
+Note that it is possible to state nonsensical worlds with conflicting
+requirements for the same part of memory. There would, however, be no
 memory that would satisfy such a world which for all intents and purposes would
@@
         \begin{aligned}
-          & \forall \ms_f, \heap', i \leq n \ldotp  (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
-          & \;\; \Rightarrow
-          \begin{aligned}[t]
-            & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
-            & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \heapSat[\hs']{n-i}{W'}
-          \end{aligned}
+          & \forall \ms_f, \heap', i \leq n \ldotp  (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap') \Rightarrow \\
+          & \;\; %\Rightarrow
+            \exists W' \futurestr W, \ms_r, \ms' \ldotp \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \heapSat[\hs']{n-i}{W'}
         \end{aligned}\right\}\\
@@
         \begin{multlined}
-          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
+          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp 
           \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W)
@@
     \left\{
-    \begin{multlined}
-      \npairP{(\var{perms}, \\
-        \start,\addrend)}
-    \end{multlined}
+      \npairP{(\var{perms}, \start,\addrend)}
      \; \middle| \;
@@
   & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
+  & \quad\;\; \text{and } \gl = \glob \Rightarrow \future = \futurestr
   \end{align*}
@@
 The logical relation defines semantically when values, program counters, and
-configurations are safe. In Section~\ref{sec:formalizing-guarantees}, we
-introduced what is meant by safe values. Similarly for program counters and
-configurations, they are safe with respect to a world when they stick to the
-memory protocols described by the world.
-
-The definition of the logical relation is found in
+configurations are safe according to the intuitions described in
+Section~\ref{sec:formalizing-guarantees}. The definition is found in
 \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
@@
 
-The \emph{observation relation} $\observations$ defines what configurations we
-consider safe. As previously states, a configuration is safe with respect to a
-world when the execution of said configuration does not break the memory
-protocols of the world. Roughly speaking, this means that when the execution of
-a configuration halts, then there is a private future world that the resulting
-memory satisfies.  Notice that failing is considered safe behavior. In fact, the
-machine often resorts to failing when an unauthorized access is attempted, like
-loading from a capability without read permission. This is similar to
-\citet{Devriese:2016ObjCap}'s logical relation for an untyped language, but
-unlike typical logical relations for typed languages, which require that
-programs do not fail.
+First, the \emph{observation relation} $\observations$ defines what
+configurations we consider safe. As previously states, a configuration is safe
+with respect to a world when the execution of said configuration does not break
+the memory protocols of the world. Roughly speaking, this means that when the
+execution of a configuration halts, then there is a private future world that
+the resulting memory satisfies. Notice that failing is considered safe behavior.
+In fact, the machine often resorts to failing when an unauthorized access is
+attempted, like loading from a capability without read permission. This is
+similar to \citet{Devriese:2016ObjCap}'s logical relation for an untyped
+language, but unlike typical logical relations for typed languages, which
+require that programs do not fail.
 
@@
 program counter if it can be plugged into a safe register file (i.e.\ a register
-file in $\stdrr$) and paired with a memory satisfying the world to become a
-safe configuration. This means that integers and non-executable capabilities
-(e.g. readonly and enter capabilities) are considered safe as program counters
+file in $\stdrr$) and paired with a memory satisfying the world to become a safe
+configuration. Note that integers and non-executable capabilities (e.g.
+$\readonly$ and $\entry$ capabilities) are considered safe program counters
 because when they are plugged into a register file and paired with a memory, the
@@
 For capabilities with read permission, the $\readCond{}$ ensures that it can
-only be used to read safe words by putting an upper bound on what can be
-required by the region governing the memory the read capability gives access
-to. An adversary can read anything the capability governs, so all the words they
-can possibly read must be safe - that is they must be in the value relation. The
-read condition also depends on the locality of the capability because, generally
-speaking, global capabilities should only depend on permanent
-regions. Concretely to make sure global capabilities only depend on permanent
-regions, we use the function $\var{localityReg}(\gl,W)$, which projects out all
-active regions when $\gl$ (the locality) is local, but only the permanent
-regions when $\gl$ is global. The upper bound on the region is enforced in terms
-of a standard region $\iota^\pwl_{\start,\addrend}$ which requires all the words
-in the range $[\start,\addrend]$ to be safe.  The region in the world that
-actually governs the part of memory the capability gives access to must as most
-be as restrictive as the standard region.  \lau{Here I would have liked to point
-  back to the bit about step-indexing in the intuition section, but
-  unfortunately the ``step down'' happens in the standard region which we have
-  not included.}
+only be used to read safe words, i.e.\ words that are in the value relation.
+This is done by imposing a lower bound on the memory requirements of the region
+governing the addressed memory. The lower bound is formulated in terms of a
+standard region $\iota^\pwl_{\start,\addrend}$, which requires all the words in
+the range $[\start,\addrend]$ to be safe, i.e.\ in the value relation. The
+region governing the addressed memory must be at least as restrictive as the
+standard region. The read condition also takes into account the locality of the
+capability because, generally speaking, global capabilities should only depend
+on permanent regions. Concretely, we use the function
+$\var{localityReg}(\gl,W)$, which projects out all active (non-revoked) regions
+when the locality $\gl$ is local, but only the permanent regions when $\gl$ is
+global. Note finally that the step-indexing described in
+Section~\ref{sec:formalizing-guarantees} is effectively at play here, but hidden
+in the definition of the standard region $\iota^\pwl_{\start,\addrend}$, which
+we omit for space reasons.
 
@@
 capability's range of authority. An adversary can use such a capability to write
-any word they can get a hold of, and we can safely assume that they can get a
-hold of any safe word, so the region governing the relevant memory must allow
-any safe word to reside there. In the name of making the logical relation as
-liberal as possible, we make this a lower bound of what the region may allow.
-\lau{Maybe add: Conceptually, the lower bound also conveys the amount of faith
-  one can put in the contents of a piece of memory that an adversary has been
-  given write access to.}  For write capabilities, we also have to take into
-account that we have two flavours of write permissions namely write and
-write-local. In the case of write-local capabilities, the region needs to at
-least require any safe word to be written, but in the case of write
-capabilities, the capability cannot be used to write local capabilties, so the
-region needs only to at least require safe non-local values.  In the write
-condition, this is handled by parameterising it with a region. For the
-write-local capabilities the write condition is applied with the standard region
-$\iota^\pwl$ and for the write capabilities we use a different standard region
-namely $\iota^\nwl_{\start,\addrend}$ which requires that the words in
-$[\start,\addrend]$ are non-local and safe. Like before, we use
+any word they can get a hold of, and we can safely assume that they can only get
+a hold of safe words, so the region governing the relevant memory must allow any
+safe word to be written there. In order to make the logical relation as liberal
+as possible, we make this a lower bound of what the region may allow. From
+another point of view, this lower bound is also an upper bound on the amount of
+trust that can be placed in the contents of memory that an adversary has been
+given write access to. For write capabilities, we also have to take into account
+the two flavours of write permissions: write and write-local. In the case of
+write-local capabilities, the region needs to allow (at least) any safe word to
+be written, but in the case of write capabilities, the capability cannot be used
+to write local capabilties, so the region only needs to allow safe non-local
+values. In the write condition, this is handled by parameterising it with a
+region. For the write-local capabilities the write condition is applied with the
+standard region $\iota^\pwl_{\start,\addrend}$ that we saw before and for the
+write capabilities we use a different standard region
+$\iota^\nwl_{\start,\addrend}$ which requires that the words in
+$[\start,\addrend]$ are non-local and safe. As before, we use
 $\var{localityReg}$ to pick an appropriate region based on the capability's
-locality. Finally the region is required to be address-stratified. Intuitively,
-this means that if a region accepts two memory segments, then it must also
-accept every memory segment ``inbetween'' that is every memory segment where
-each address contains a value from one of the two memory segments. An
-interesting property of the read and write condition is that they prohibit
-global write-local capabilities which, as discussed in
-Section~\ref{sec:stack-and-return-pointer}, is necessary for the safe use of
+locality. Finally, there is a technical requirement that the region must be
+\emph{address-stratified}. Intuitively, this means that if a region accepts two
+memory segments, then it must also accept every memory segment ``in between'',
+that is every memory segment where each address contains a value from one of the
+two memory segments. An interesting property of the write condition is that they
+prohibit global write-local capabilities which, as discussed in
+Section~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
 local capabilities.
@@
 address they point to. They also change permission when jumped to, so we require
-them to be safe as a program counter when the permission is changed to $\exec$.
+them to be safe as a program counter after the permission is changed to $\exec$.
 Because the capabilities are not necessarily invoked immediately, this must be
-true in any future world, but it depends on the locality of the capability which
-future worlds we consider. If it is global, then it should only depend on
-permanent regions, so we require safety as a program counter in \emph{private}
-future worlds (recall that temporary regions may be revoked in a private future
-world). For local capabilities, it suffices to be safe in \emph{public} future
+true in any future world, but it depends on the capability's locality which
+future worlds we consider. If it is global, then we require safety as a program
+counter in \emph{private} future worlds (where temporary regions may be
+revoked). For local capabilities, it suffices to be safe in \emph{public} future
 worlds, where temporary regions are still present.
@@
 Section~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
-capabilities that points to arbitrary instructions. The relation between
+capabilities that point to arbitrary instructions. The relation between
 Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow
@@
 there are many details to the formal treatment, and therefore we
-necessarily have to omit some details in the lemmas we state here. The
+necessarily omit some details in the lemma statements. The
 examples may look deceivingly short, but it is because they use the
@@
 that the configuration just after \texttt{\footnotesize{scall}} is done cleaning
-up is safe. We use the Fundamental Theorem to reason about the unknown adversarial code, but
-notice that the adversary capability is an enter capability, which the
-Fundamental Theorem says nothing about.
-\dominique{why doesn't it?}
-\lau{The FTLR says nothing about enter capabilities, because an enter capability cannot directly be used as a program counter.
-The reason being that the computation would fail immediately and the condition in the observation relation would immediately be true (as it requires the computation to terminate and not fail).
-It would be trivial to show that the enter capability, data and other capabilities without execute permission is in the expression relation.}
-Luckily the enter capability becomes
-$\exec$ after the jump and then the Fundamental Theorem applies.
+up is safe. We use the Fundamental Theorem to reason about the unknown
+adversarial code, but notice that the adversary capability is an enter
+capability, which the Fundamental Theorem says nothing about. Luckily the enter
+capability becomes $\exec$ after the jump and then the Fundamental Theorem
+applies.
 
@@
 The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}},
-is a faithful translation of a tricky example  known from the
-literature (known as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
-parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
+is a faithful translation of a tricky example known from the literature (known
+as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists
+of two parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
 \texttt{\footnotesize{g1}} is a closure generator that generates closures with
 one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
-the program (note we can omit some calling convention security measures
-because the stack is not used or passed to the adversary in the closure
-generator). \texttt{\footnotesize{f4}} expects one argument, a callback. It sets
-$x$ to $0$ and calls the callback. When it returns, it sets $x$ to $1$ and calls
-the callback a second time. When it returns again, it asserts $x$ is $1$
-and returns. This example is more complicated than the previous ones because it
-involves a closure invoked by the adversary and an adversary callback invoked by
-us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means that we
-need to check (1) that the stack pointer that the closure receives from the
-adversary has write-local permission and (2) that the adversary callback is global.
+the program (note we can omit some calling convention security measures because
+the stack is not used in the closure generator). \texttt{\footnotesize{f4}}
+expects one argument, a callback. It sets $x$ to $0$ and calls the callback.
+When it returns, it sets $x$ to $1$ and calls the callback a second time. When
+it returns again, it asserts $x$ is $1$ and returns. This example is more
+complicated than the previous ones because it involves a closure invoked by the
+adversary and an adversary callback invoked by us. As explained in
+Section~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
+that the stack pointer that the closure receives from the adversary has
+write-local permission and (2) that the adversary callback is global.
 
@@
 
-In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in the
-proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
-all the situations where control is passed to the adversary. The potentially
-local return pointer received by \texttt{\footnotesize{f4}} from the adversary is safe in
-$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
-constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
-by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
-we pass away and static temporary regions (a region that accepts exactly one
-memory) for our local stack as well as the adversary's local stack. The given
-worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
-to return to us. None of these worlds can have masked the temporary regions of
-$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
-regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
-$W_6$ a public future world of $W_1$. It is therefore safe to use the return
-pointer to return to the adversary, since we have restored validity of any local
-state they might have stashed away.
+In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
+the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
+for all the situations where control is passed to the adversary. The potentially
+local return pointer received by \texttt{\footnotesize{f4}} from the adversary
+is safe in $W_1$, so it can only be used in public future worlds. Let us see how
+$W_6$ is constructed as a public future world of $W_1$, as well as a private
+future world of $W_5$. The worlds $W_2$ and $W_4$ are constructed by revoking
+all temporary regions and adding a $\iota^\pwl$ region for the stack we pass
+away and static temporary regions (a region that accepts exactly one memory) for
+our local stack as well as the adversary's local stack. The given worlds $W_3$
+and $W_5$ are public future worlds in which the adversary chooses to return to
+us. None of these worlds can have masked the temporary regions of $W_1$ with
+permanent regions. $W_6$ is constructed by revoking the temporary regions of
+$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$ a public
+future world of $W_1$. It is therefore safe to use the return pointer to return
+to the adversary, since we have restored validity of any local state they might
+have stashed away.
 
@@
   \node[main node] (1) {$W_1$};
-  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f3} called}}};
+  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f4} called}}};
   \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};
@@
   \node[main node] (6) [above right of=5]{$W_6$};
-  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
+  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f4} returns}}};
   \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
@@
 
-\emph{Modularity} It is important that our calling convention is modular, i.e.
-we do not assume that our code is specially privileged w.r.t. the adversary, and
+\emph{Modularity} It is important that our calling convention is modular, i.e.\ 
+we do not assume that our code is specially privileged w.r.t.\ the adversary, and
 they can apply the same measures to protect themselves from us as we do to
@@
 Local capabilities are a central part of the calling convention as they are used
-to construct stack and return pointer. The use of local capabilities for the
-calling convention unfortunately limits the extend of which local capabilties
+to construct stack and return pointers. The use of local capabilities for the
+calling convention unfortunately limits the extent to which local capabilties
 can be used for other things. Say we are using the calling convention and
@@
 pointer. That is, if we first push something to the stack and then write to the
-local capability, then we may overwrite our local state. While this sounds
-unsafe, the safety actually depends on what our memory invariants we want to
-establish, so the safety is actually dictated by the logical relation. This may,
-however, not be a satisfactory requirement for say compiler developers as it is
-impossible to figure out what local capabilities are safe to use. A different
-approach would be to do dynamic checks on local capabilities, but this may be to
-costly to do in practice.
+local capability, then we may be (tricked into) overwriting our own local state.
+The logical relation helps by telling us what we need to ascertain or check in
+such scenario's to guarantee safety and preserve our invariants, but such checks
+may be costly and it is not clear to us whether there are practical scenario's
+where this might be realistic.
 
@@
 We also need to be careful when we receive a capability from an adversary that
-we want to pass on to a different adversary. It turns out that the logical
-relation again dictates when this is safe. Namely, the logical relation says
-that we can only pass on safe arguments. For instance, when we receive a stack
-pointer from an adversary, then we may at some point want to pass on part of
-this stack pointer to say a callback. In order to do so, we need to make sure
-the stack pointer is safe which means that it must only point to safe
-values. When received from an adversary, we have to consider the contents of the
-stack unsafe, so before we pass it on we clear which corresponds to filling it
-with safe values as integers are always safe to pass on. In general, if we want
-to pass on a local capability we received from an adversary, we have to make
-sure it is safe, for instance by clearing everything it points to (if possible,
-the capability might be a readonly capability). Clearing everthing is, however,
-not always a desirable approach as we may have an array or data structure with
-important values we want to pass on by reference in which case it is
-unacceptable to just clear the entire thing. In this case, it might be possible
-to make a deep safety analysis to make sure that the capability points to
-nothing dangerous. What we are left with is a too conservative approach or an
-non-obvious expensive dynamic safety analysis.
+we want to pass on to a different (instance of the) adversary. It turns out that
+the logical relation again tells us when this is safe. Namely, the logical
+relation says that we can only pass on safe arguments. For instance, when we
+receive a stack pointer from an adversary, then we may at some point want to
+pass on part of this stack pointer to, say, a callback. In order to do so, we
+need to make sure the stack pointer is safe which means that, if we have revoked
+temporary invariants, the stack must not directly or indirectly allow access to
+local values that we cannot guarantee safety of. When received from an
+adversary, we have to consider the contents of the stack unsafe, so before we
+pass it on, we have to clear it, or perform a dynamic safety analysis of the
+stack contents and anything it points to. Clearing everything is not always
+desirable and a dynamic safety analysis is hard to get right and potentially
+expensive.
 
@@
 %Not unheard of - other machine also have things specific for language features.
-The use of local capabilities for other things than stack and return pointer
-turns out to be somewhat limitted. It is, however, not unusual that machines
-have builtin special constructs that are used for just one language feature. A
-machine could for instance have call and return instructions with the sole
-prupose of handling everything related to calls.
+In summary, the use of local capabilities for other things than stack and return
+pointers is likely only possible in very specific scenario's when using our
+calling convention. While this is unfortunate, it is not unheard of that
+processors have builtin constructs that are exclusively used for handling
+control flow, such as, for example, the call and return instructions that exist
+in some instruction sets.
 \\\\
@@
 \emph{Single stack} A single stack is a good choice for the simple capability
-machine presented here, because it works well with higher-order functions. The
-alternative to a single stack would be to have one stack per component. With
-multiple stacks on this simple capability machine, it is not entirely obvious
-how each component would retrieve their stack pointer. Further with higher-order
-functions, it is not clear which stack the higher-order function should
-use. Should it use the stack of the component that created it or the stack of
-the component in which it is called. If we let it use the stack of the component
-that uses the higher-order function, then what is the purpose of having multiple
-stacks when the stack is not exclusive to each component. If we let it use the
-stack of its creator, then the question is how do you obtain this stack
-pointer. We do not claim that it is impossible to make it work with multiple
-stack in the context of higher-order functions, but it is not obvious how it
-would work.
+machine presented here, because it works well with higher-order functions. An
+alternative to a single stack would be to have a separate stack per component.
+The trouble with this approach is that, with multiple stacks and local stack
+pointers, it is not clear how components would retrieve their stack pointer upon
+invocation without compromising safety. A safe approach could be to have stack
+pointers stored by a central, trusted stack management component, but it is not
+clear how that could scale to large numbers of separate components. Handling
+large numbers of components is a requirement if we want to use
+capability machines to enforce encapsulation of, for example, every object in an
+object-oriented program or every closure in a functional program.
 
@@
 that will be preserved when we do, but it does not tell us much about the
-address of the capability.
-
-To make this very concrete, consider how capabilities with permission $\noperm$
-are always in the value relation, irrespective of their properties.
-Semantically, this makes perfect sense, since they are always safe since they
-cannot be used for anything. However, it also means we do not get syntactic
-information about them.
+address of the capability. As a very concrete example, capabilities with
+permission $\noperm$ are always in the value relation, irrespective of their
+address. Semantically, this makes perfect sense, since they are always safe
+since they cannot be used for anything anyway. However, it also means we do not
+get syntactic information about them.
 
@@
 However, in other situations, we may be interested in proving more syntactic
-properties like the more syntactic properties, that are often considered in
-object capability literature like confinement, no authority amplification etc.
-Although such properties are typically more restrictive and less easy to use for
-reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation
-much like ours can be adapted to also support them, by quantifying the logical
-relation over a custom interpretation of effectful computations and the type of
+properties like the ones that are often considered in object capability
+literature: confinement, no authority amplification etc. Although such
+properties are more restrictive and less easy to use for reasoning,
+\citet{Devriese:2016ObjCap} have demonstrated how a logical relation like
+ours can be adapted to also support them, by quantifying the logical relation
+over a custom interpretation of effectful computations and the type of
 references. We expect their solution can be readily adapted to our setting,
 modulo some details (like the fact that we do not just have read-write
-capabilities, but also other types).\\\\
+capabilities, but also others).\\\\
 % \emph{No program logic} TODO: discuss this?
@@
 plugged with safe values. This is natural because we model arbitrary assembly
-code that does not necessarily respect a particular calling convention. As a
-result, return pointers are in principle values like all others and there is no
-reason to treat them specially in the logical relation.
+code that does not necessarily respect a particular calling convention: return
+pointers are in principle values like all others and there is no reason to treat
+them specially in the logical relation.
 
@@
 secure compilation
-result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
+result \citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
 differs from theirs in that we use a different form of low-level security
@@
 with recursive
-worlds~\citep{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
+worlds \citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
 closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
@@
 relation uses several ideas from previous work, like Kripke worlds with regions
-containing state transition systems~\citep{Ahmed:popl09}, public/private future
-worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
-biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
+containing state transition systems \citep{Ahmed:popl09}, public/private future
+worlds \citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
+biorthogonality \citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
 Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
-object capability patterns\citep{swasey:2017}. The logic is based on
-Iris~\citep{iris,iris2,iris3}, a state of the art higher-order
+object capability patterns \citep{swasey:2017}. The logic is based on
+Iris \citep{iris,iris2,iris3}, a state of the art higher-order
 concurrent separation logic and is formalized in Coq, building on the
-Iris Proof Mode for Coq~\citep{ipm}.  
-XXX HERE
+Iris Proof Mode for Coq \citep{ipm}.
 

### 2017-06-23 popl-2018/paper.tex

@@
 
-Our logical relation is a unary, step-indexed Kripke logical relation
-with recursive
-worlds \citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
+Our logical relation is a unary, step-indexed Kripke logical relation with
+recursive worlds
+\citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
 closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
 capability safety in a high-level JavaScript-like lambda calculus. Our
-Fundamental Theorem is similar to theirs and expresses
-capability safety of the capability machine. Because we are not interested in
-externally observable side-effects (like console output or memory access
-traces), we do not require their notion of effect parametricity. Our logical
-relation uses several ideas from previous work, like Kripke worlds with regions
-containing state transition systems \citep{Ahmed:popl09}, public/private future
-worlds \citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion), and
-biorthogonality \citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
-Swasey et. al.\ have recently developed a \emph{logic}, OCPL,  for verification of
-object capability patterns \citep{swasey:2017}. The logic is based on
-Iris \citep{iris,iris2,iris3}, a state of the art higher-order
-concurrent separation logic and is formalized in Coq, building on the
-Iris Proof Mode for Coq \citep{ipm}.
+Fundamental Theorem is similar to theirs and expresses capability safety of the
+capability machine. Because we are not interested in externally observable
+side-effects (like console output or memory access traces), we do not require
+their notion of effect parametricity. Our logical relation uses several ideas
+from previous work, like Kripke worlds with regions containing state transition
+systems \citep{Ahmed:popl09}, public/private future worlds \citep{Dreyer:jfp12}
+(see Section~\ref{sec:discussion} for a discussion), and biorthogonality
+\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
+
+Swasey et. al.\ have recently developed a \emph{logic}, OCPL, for verification
+of object capability patterns~\citep{swasey:2017}. The logic is based on
+Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
+separation logic and is formalized in Coq, building on the Iris Proof Mode for
+Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
+capability safety for a lambda-calculus (with concurrency) compared to the
+earlier work by \citet{Devriese:2016ObjCap}. In the future we would also like to
+investigate a new program logic for reasoning about capability safety for our
+capability machine model. We think Iris would also be a natural starting point
+for such an endeavour, since Iris is really a framework, which can be
+instantiated to different programming languages. OCPL was able to leverage
+existing Iris specifications for the high-level language; for our capability
+machine model, however, it would be necessary to devise new kinds of
+specifications for our low-level programs with unstructured control-flow. It is
+likely that we could get inspiration from earlier work on logics for assembly
+programming languages, such as XCAP~\citep{xcap}.
 

### 2017-06-23 popl-2018/paper.tex
Add a statement of the double monotonicity of our value relation.
Slight rephrasing.
Remove an outdated TODO note.

@@
 local-state encapsulation.  
-\lau{Do we say anything about arguments to the callback?
+% \lau{Do we say anything about arguments to the callback?
 Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary.
 Can we even to any extend allow local arguments that we pass on from an unknown source?}
+% Domi: this is now discussed in the discussion section, no need to discuss it here?
 
@@
 
+In the technical appendix, we prove that (as explained before), safety of all
+values is preserved in public future worlds, but safety of global values is
+also preserved in private future worlds:
+\begin{lemma}[Double monotonicity of value relation]~
+  \begin{itemize}
+  \item If $W' \futurewk W$ and $\npair{w} \in \stdvr(W)$, then $\npair{w} \in
+    \stdvr(W')$.
+  \item If $W' \futurestr W$ and $\npair{w} \in \stdvr(W)$ and $w =
+    \stdcap[(\perm,\glob)]$ (i.e.\ $w$ is a global capability), then $\npair{w}
+    \in \stdvr(W')$.
+  \end{itemize}
+\end{lemma}
+
 \subsection{Safety of the capability machine}

### 2017-06-24 fully-abstract-comp/technical-report.tex
merge

@@
 \newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
+\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
 
@@
 \newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
+\newcommand{\scseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}
 
@@
 \newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
-\newcommand{\tcseal}[2]{\twoinstr{\trg{cseal}}{#1}{#2}}
 
@@
 \newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}
+\newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}
 
@@
 \newcommand{\opc}{\var{opc}}
+\newcommand{\vsc}{\var{sc}}
+\newcommand{\cb}{\var{cb}}
 
@@
 \newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
-\newcommand{\clearms}[1]{\plainfun{clear_{ms}}{#1}}
-\newcommand{\clearlocreg}[1]{\plainfun{clear_{loc,reg}}{#1}}
+\newcommand{\nonexec}[1]{\plainfun{nonExecutable}{#1}}
+\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
+\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
 
@@
       & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
-      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}
+      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
 \end{array}
@@
 \[
-  \updPcAddr \defeq \dots
+  \updPcAddr{} \defeq \dots
 \]
@@
 \[
-  \updPcPerm \defeq \dots
+  \updPcPerm{} \defeq \dots
 \]
@@
 \[
-  \clearms \defeq \dots
+  \nonexec{} \defeq \dots
 \]
@@
 \[
-  \clearlocreg \defeq \dots
+  \noRetStkReg{} \defeq \dots %remember to ignore certain registers
+\]
+
+\[
+  \noRetStkMs{} \defeq \dots %remember to ignore certain registers
 \]
@@
 \begin{align*}
-  \ssem{\sjmp{r}} = & 
-    \begin{cases}
-      \Phi
-      \begin{aligned}[t]
-        & \update{\reg.\pcreg}{\opc} \\
-        & \update{\smsstk}{\ms_\stk' \uplus \clearms{\ms_\stk}}\\
-        & \update{\sastk}{a_\stk'}\\
-        & \update{\sstk}{\stk'}
-      \end{aligned} & \Phi(r) = \retptr \text{ and }
-      \Phi.\sstk = (\opc,\ms_\stk',a_\stk') :: \stk'\\
-      & \\
-      \updPc{\updPcPerm{\Phi(r)}} & \text{otherwise}
-    \end{cases} \\
-\end{align*}
-\begin{align*}
   \ssem{\scall{r_1}{r_2}} = & 
     \begin{cases}
-      (\mem,\reg',\stk',\ms_\stk') & \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and }\\
+      (\mem,\reg',\stk',\ms_\stk') & \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \dots\\
       \\
@@
 \end{align*}
+
+The $\scseal{}{}{}$ instruction seals a sealable capability with a seal.
+\begin{align*}
+  \ssem{\scseal{r_1}{r_2}{r_3}} = & 
+                                    \begin{cases}
+                                      \updReg{r_1}{\vsc} & 
+                                      \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        \Phi(r_2) = c \in \SealableCaps \text{ and } \\
+                                        \Phi(r_3) = \seal{\sigma,\_} \text{ and } \\
+                                        \vsc = \sealed{c,\sigma}
+                                      \end{array}\\
+                                      & \\
+                                      \failed & \text{otherwise}
+                                    \end{cases}
+\end{align*}
+
+Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$.
+We cannot just use $\sjmp{}$ because it only takes one argument and we need to be able to jump to a code/data pair.
+To this end we introduce $\sxjmp{}{}$:
+\begin{align*}
+  \ssem{\sxjmp{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \Phi
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  [\reg.\pcreg \mapsto c_1] \\
+                                  [\reg.\rdata \mapsto c_2]
+                                \end{array} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
+                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
+                                  \nonexec{c_2}
+                                \end{array}\\
+%
+                                \\
+%
+                                \failed & \text{otherwise}
+                              \end{cases}
+\end{align*}
+\lau{This is $\sxjmp{}{}$ for the target language - this one also needs to allow jumps to the return pointer}
+Notice, $\rdata$ serves the same function as the IDC register in CHERI, i.e.\ this is where the data capability is put.
+In the interpretation of $\sxjmp{}{}$, we require $c_2$ to be non-executable because CHERI does it.
+We are, however, not certain whether this requirement is strictly necessary (maybe it is dangerous to $\sxjmp{r}{r}$ where $r$ points to the same capability?
+Maybe it can be use to unseal a code capability by first sealing another code capability with the seal, $\sxjmp{}{}$ with the newly sealed and the old sealed capability.
+Now both capabilities are unsealed).
+CHERI also checks whether $c_1$ is executable and in range, but we do this on every step of the operational semantics, so we do not repeat that check.
+
+\begin{align*}
+  \ssem{\sjmp{r}} = & \dots
+\end{align*}
+
 \begin{itemize}
@@
 \subsubsection{Target Language}
+
+\begin{align*}
+  \ssem{\sxjmp{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \Phi
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  [\reg.\pcreg \mapsto c_1] \\
+                                  [\reg.\rdata \mapsto c_2]
+                                \end{array} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
+                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
+                                  \nonexec{c_2}
+                                \end{array}
+\\
+\failed & \text{otherwise}
+                              \end{cases}
+\end{align*}
 

### 2017-06-29 popl-2018/paper.tex
merge
Added moderation of feasibillity of implementation of instruction.

@@
 %% For double-blind review submission
-\documentclass[anonymous,acmsmall,review]{acmart}\settopmatter{printfolios=true}
+\documentclass[acmsmall,review,10pt,anonymous]{acmart}\settopmatter{printfolios=true}
 %% For single-blind review submission
-%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
 %% For final camera-ready submission
-%\documentclass[acmlarge]{acmart}\settopmatter{}
+%\documentclass[acmsmall]{acmart}\settopmatter{}
 
@@
 Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
-calling convention without security guarantees. However, most of our security measures
-require only a few atomic checks or register clearings on boundary crossings
-between trusted code and adversary, which should produce an acceptable
+calling convention without security guarantees. However, most of our security
+measures require only a few atomic checks or register clearings on boundary
+crossings between trusted code and adversary, which should produce an acceptable
 performance overhead. The only exception are the requirements for stack clearing
@@
 special support for this requirement, in the form of a highly-optimized
-instruction for erasing a large block of memory.
+instruction for erasing a large block of memory. Nevertheless, from a discussion
+with the designers of the CHERI capability machine, we gather that it is not immediately
+clear whether and how such a primitive could be implemented efficiently in the CHERI context.
 

### 2017-06-29 popl-2018/paper.tex
commented out some boilerplate instructions and disabled our comments.

@@
 % Uncomment me to disable comments
-% \renewcommand\lau[1]{}
-% \renewcommand\dominique[1]{}
-% \renewcommand\lars[1]{}
+\renewcommand\lau[1]{}
+\renewcommand\dominique[1]{}
+\renewcommand\lars[1]{}
 
@@
 %% Supplied to authors by publisher for camera-ready submission
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{1}
-\acmArticle{1}
-\acmYear{2017}
-\acmMonth{1}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\else\makeatother
+% \acmJournal{PACMPL}
+% \acmVolume{1}
+% \acmNumber{1}
+% \acmArticle{1}
+% \acmYear{2017}
+% \acmMonth{1}
+% \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+% \startPage{1}
+% \else\makeatother
 %% Conference information (used by SIGPLAN proceedings format)
 %% Supplied to authors by publisher for camera-ready submission
-\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
-\acmYear{2017}
-\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\fi
+% \acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
+% \acmYear{2017}
+% \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+% \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+% \startPage{1}
+% \fi
 
@@
 %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+% \begin{CCSXML}
+% <ccs2012>
+% <concept>
+% <concept_id>10011007.10011006.10011008</concept_id>
+% <concept_desc>Software and its engineering~General programming languages</concept_desc>
+% <concept_significance>500</concept_significance>
+% </concept>
+% <concept>
+% <concept_id>10003456.10003457.10003521.10003525</concept_id>
+% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
+% <concept_significance>300</concept_significance>
+% </concept>
+% </ccs2012>
+% \end{CCSXML}
+
+% \ccsdesc[500]{Software and its engineering~General programming languages}
+% \ccsdesc[300]{Social and professional topics~History of programming languages}
 %% End of generated code
@@
 in some instruction sets.
-\\\\
 % Old version of the above:
@@
 object-oriented program or every closure in a functional program.
-
+\\\\
 \noindent \textbf{Reasoning about capability machine programs}\\
@@
 modulo some details (like the fact that we do not just have read-write
-capabilities, but also others).\\\\
+capabilities, but also others).
 % \emph{No program logic} TODO: discuss this? 
-
+\\\\
 \noindent\textbf{Logical relation}\\

### 2017-06-29 popl-2018/paper.tex
typos and bad cite

@@
 However, there is one exception to the rule above: local capabilities can be
-stored to memory through a capability with write-local authority
+stored to memory for which we have a capability with write-local authority
 (i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
@@
 safe). Concretely, worlds may contain \emph{temporary} regions, representing
-invariants that only local capabilties may rely on for their safety, and which
+invariants that only local capabilities may rely on for their safety, and which
 may be revoked (disabled) in private future worlds.
@@
 can be revoked in private future worlds. Doing this may break the safety of
-local capbalities but not global ones. This means that local capabilities
+local capabilities but not global ones. This means that local capabilities
 can safely depend on the protocols imposed by temporary regions, but global
@@
 be written, but in the case of write capabilities, the capability cannot be used
-to write local capabilties, so the region only needs to allow safe non-local
-values. In the write condition, this is handled by parameterising it with a
+to write local capabilities, so the region only needs to allow safe non-local
+values. In the write condition, this is handled by parameterizing it with a
 region. For the write-local capabilities the write condition is applied with the
@@
 to construct stack and return pointers. The use of local capabilities for the
-calling convention unfortunately limits the extent to which local capabilties
+calling convention unfortunately limits the extent to which local capabilities
 can be used for other things. Say we are using the calling convention and
@@
 calling convention. While this is unfortunate, it is not unheard of that
-processors have builtin constructs that are exclusively used for handling
+processors have built-in constructs that are exclusively used for handling
 control flow, such as, for example, the call and return instructions that exist
@@
 
-Swasey et. al.\ have recently developed a \emph{logic}, OCPL, for verification
-of object capability patterns~\citep{swasey:2017}. The logic is based on
+\citet{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
+of object capability patterns. The logic is based on
 Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent

### 2017-07-03 popl-2018/paper.tex
small updates/typos

@@
   them an interesting target for compilation schemes that provably enforce
-  properties like control-flow correctness, encapsulation of local state. We
+  properties like control-flow correctness and encapsulation of local state. We
   provide a formalization of a representative capability machine with local
@@
 load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and arithmetic addition operators ($\lt{}{}{}$, $\plus{}{}{}$ and
+and $\store{}{}$) and arithmetic addition operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and
 $\minus{}{}{}$, operating only on numbers). There are three instructions for
@@
 data through the updated version of $c$ that is then in $\pcreg$\footnote{Note
-  that the capability in the $\pcreg$-register can be used to load from.}.
-% \lau{As we
-%   talked about what is described here, is not how we make closures in the TR, so
-%   we should consider whether we should refrain from using that word here. }
+that the capability in the $\pcreg$-register can be used to load from.}.  %
+\lau{As we % talked about what is described here, is not how we make closures in
+the TR, so % we should consider whether we should refrain from using that word
+here. }
 
@@
 (i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
-where register contents can be stored, including local capabilities. As long as all
-capabilities with write-local authority are themselves local and the stack is
-cleared after control is passed back by the adversary, we will see that this
+where register contents can be stored, including local capabilities. As long as
+all capabilities with write-local authority are themselves local and the stack
+is cleared after control is passed back by the adversary, we will see that this
 does not break the intended behavior of local capabilities.
@@
 
-% Contrary to more traditional assembly languages, capability machines do not
-% allow literal addresses to appear as part of instructions. This means that we
+% Contrary to more traditional assembly languages, capability machines do not %
+allow literal addresses to appear as part of instructions. This means that we
 
@@
 \section{Stack and return pointer management using local capabilities}
-\label{sec:stack-and-return-pointer}
-One of the contributions in this paper is a demonstration that local
-capabilities on a capability machine support a calling convention that enforces
-control-flow correctness in a way that is provably watertight, potentially
-efficient, does not rely on a trusted central stack manager, and supports
-higher-order interfaces to an adversary where an adversary is just some unknown piece of code. In this section, we explain this
+\label{sec:stack-and-return-pointer} One of the contributions in this paper is a
+demonstration that local capabilities on a capability machine support a calling
+convention that enforces control-flow correctness in a way that is provably
+watertight, potentially efficient, does not rely on a trusted central stack
+manager and supports higher-order interfaces to an adversary, where an adversary
+is just some unknown piece of code. In this section, we explain this
 convention's high-level approach, the security measures to be taken in a number
@@
 
-In this section, we construct such a formalisation. We make use of some
+In this section, we construct such a formalization. We make use of some
 well-studied and powerful (but non-trivial) machinery from the literature.
@@
 \paragraph{Local capabilities}
-A final thing we explain here is how our results take into account local
+A final thing we provide informal intuition for is how our results take into account local
 capabilities and their special treatment by the hardware. Normally, when we
@@
 can allow the second invocation to happen in \emph{private} future worlds ($W'
-\futurestr W$), in which global safe capabilities remain safe, but local ones
+\futurestr W$), in which safe global capabilities remain safe, but local ones
 don't. This private future world relation is more liberal than the standard,
@@
 the temporary and permanent regions is a state interpretation function that
-determines what memory segments the region permits in each state of its states.
-The different monotonicity requirements in the two interpretation functions
-reflects how permanent regions rely only on permanent protocols whereas
-temporary regions can rely on both temporary and permanent protocols.
-$\UPred{\HeapSegments}$ is a step-indexed, downwards closed predicate on memory
-segments.
+determines what memory segments the region permits in each state of the state
+transition system.  The different monotonicity requirements in the two
+interpretation functions reflects how permanent regions rely only on permanent
+protocols whereas temporary regions can rely on both temporary and permanent
+protocols.  $\UPred{\HeapSegments}$ is a step-indexed, downwards closed
+predicate on memory segments.
 
@@
 First, the \emph{observation relation} $\observations$ defines what
-configurations we consider safe. As previously states, a configuration is safe
+configurations we consider safe. As previously stated, a configuration is safe
 with respect to a world when the execution of said configuration does not break
@@
 safe. In \figurename~\ref{fig:perm-conds}, we define such a condition for each
-kind of permission a capability can have (we will call these permission-based
-conditions).
+kind of permission a capability can have.
 
 For capabilities with read permission, the $\readCond{}$ ensures that it can
-only be used to read safe words, i.e.\ words that are in the value relation.
-This is done by imposing a lower bound on the memory requirements of the region
-governing the addressed memory. The lower bound is formulated in terms of a
+only be used to read safe words, i.e.\ words in the value relation.
+This is done by imposing an upper bound on the memory requirements of the region
+governing the addressed memory. The upper bound is formulated in terms of a
 standard region $\iota^\pwl_{\start,\addrend}$, which requires all the words in
@@
 region. For the write-local capabilities the write condition is applied with the
-standard region $\iota^\pwl_{\start,\addrend}$ that we saw before and for the
-write capabilities we use a different standard region
+standard region $\iota^\pwl_{\start,\addrend}$ that we described previously. For
+the write capabilities we use a different standard region
 $\iota^\nwl_{\start,\addrend}$ which requires that the words in
@@
 that is every memory segment where each address contains a value from one of the
-two memory segments. An interesting property of the write condition is that they
-prohibit global write-local capabilities which, as discussed in
+two accepted memory segments. An interesting property of the write condition is
+that they prohibit global write-local capabilities which, as discussed in
 Section~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of

### 2017-08-18 fully-abstract-comp/technical-report.tex
Added lt instruction and fixed minor things

@@
 \newcommand{\ssubseg}[3]{\threeinstr{\src{subseg}}{#1}{#2}{#3}}
+\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
 \newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
@@
 \newcommand{\tsubseg}[3]{\threeinstr{\trg{subseg}}{#1}{#2}{#3}}
+\newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
 \newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
@@
 \subsection{Syntax}
-The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle} (Inspired by CHERI).
+The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (Inspired by CHERI).
 The syntax of the instructions of the target machine is defined as follows:
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
       & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
@@
 \subsection{Permissions}
-\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.}
+\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.} % I am not sure why I did not think the opaque capabilities where necessary, but I have a feeling we talked about this in Leuven and came to the conclusion that they were necessary.
 \begin{figure}[!h]
@@
   \end{itemize}
-\item Considered PORs for transition systems to: (a->b => a=a and b=b)
+\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a = a \text{ and } b = b$)
   \begin{itemize}

### 2017-08-21 fully-abstract-comp/technical-report.tex
updated call semantics
Added some description and formatted the definition a little bit better

@@
 \newcommand{\ssem}[2][\Phi]{\sem{#2}_{\mathrm{src}}(#1)}
+\newcommand{\dom}{\mathrm{dom}}
 
@@
 \newcommand{\failed}{\mathrm{failed}}
+\newcommand{\undefined}{\mathrm{undefined}}
 \newcommand{\halted}{\mathrm{halted}}
@@
 \newcommand{\stk}{\var{stk}}
+\newcommand{\ret}{\var{ret}}
+\newcommand{\data}{\var{data}}
+\newcommand{\priv}{\var{priv}}
 \newcommand{\opc}{\var{opc}}
@@
 \newcommand{\cb}{\var{cb}}
+\newcommand{\baddr}{\var{b}}
+\newcommand{\eaddr}{\var{e}}
+\newcommand{\aaddr}{\var{a}}
+\newcommand{\stdrng}{[\baddr,\eaddr]}
+
+%%% Constants
+\newcommand{\constant}[1]{\mathrm{#1}}
+\newcommand{\calllen}{\constant{call\_len}}
 
@@
 \newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
-\newcommand{\nonexec}[1]{\plainfun{nonExecutable}{#1}}
+\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
+\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
 \newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
@@
     &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
-    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\     &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
+    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\     &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptr}\\ 
     &\RegName & \defbnf &\pcreg \mid \rO \mid \rstk \mid \rdata \mid \dots \\
@@
 \trg{r} &\in &  \tRegName \\
-\trg{\rn} &::= &  \trg{r} \mid n \\
-\trg{i} &::= & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
+\trg{\rn} &\defbnf &  \trg{r} \mid n \\
+\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
@@
     \src{r} &\in &  \RegName \setminus \{\rO,\rstk\}\\
-    \src{\rn} &::= & \src{r} \mid n \\
-    \src{i} & ::= &  \trg{i} \mid \scall{\src{r}}{\src{r}}
+    \src{\rn} &\defbnf & \src{r} \mid n \\
+    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
   \end{array}
@@
 \[
-  \nonexec{} \defeq \dots
+  \nonExec{\cb} \defeq \dots
 \]
@@
 \[
-  \noRetStkReg{} \defeq \dots %remember to ignore certain registers
+  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rO, \rdata \} \ldotp \reg(r) \neq \stkptr(\_)
 \]
@@
 \[
-  \noRetStkMs{} \defeq \dots %remember to ignore certain registers
+  \noRetStkMs{\ms} \defeq \defeq \forall \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) \neq \stkptr(\_)
 \]
@@
 \lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update.}
+\textbf{Call}\\
+Call has three cases, a successful call, undefined and failed.
+For a call to be successful, it needs two sealed capabilities - a data and a code capability.
+The code capability must be non-executable and the stack register must contain a stack pointer.
+Futher the code and data capability must be local and there must be no attempt at passing on the stack capability.
+\lau{Add more description here.}
+
+For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}
+
+If we are not in one of the two cases above, then the execution fails.
 \begin{align*}
@@
     \begin{cases}
-      (\mem,\reg',\stk',\ms_\stk') & \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \dots\\
+      (\mem,\reg',\stk',\ms_\stk') & \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        % "red conditions" (reference to blackboard picture)
+                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
+                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
+                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
+                                        \nonExec{cb_2} \text{ and } \\
+                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
+                                        % "yellow conditions"
+                                        \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
+                                        \noRetStkReg{\reg} \text{ and } \\
+                                        \noRetStkMs{\ms_\stk'} \text{ and } \\
+                                        \nonLoc{\cb_1,\cb_2} \text{ and } \\
+                                        \reg' = \reg
+                                        \begin{array}[t]{l}
+                                        \update{\pcreg}{\cb_1}\update{\rstk}{\stkptr{\perm,[\aaddr,eaddr],\aaddr}}\\
+                                        \update{\rO}{\retptr}\update{\rdata}{\cb_2} \text{ and }
+                                        \end{array}
  \\
+                                        \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
+                                        \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
+                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
+                                        \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
+                                      \end{array}\\
+                                      & \\
+      \undefined & \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        % "red conditions" (reference to blackboard picture)
+                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
+                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
+                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
+                                        \nonExec{cb_2} \text{ and } \\
+                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
+                                        % "yellow conditions"
+                                        \text{not } \arraycolsep=0pt
+                                        \left(\begin{array}{l}
+                                          \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
+                                          \noRetStkReg{\reg} \text{ and } \\
+                                          \noRetStkMs{\ms_\stk'} \text{ and } \\
+                                          \nonLoc{\cb_1,\cb_2}
+                                        \end{array} \right)
+                                      \end{array}\\
+                                      & \\
       \failed & \text{otherwise}
@@
                                   \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
-                                  \nonexec{c_2}
+                                  \nonExec{c_2}
                                 \end{array}\\
@@
                                   \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
-                                  \nonexec{c_2}
+                                  \nonExec{c_2}
                                 \end{array}

### 2017-08-21 fully-abstract-comp/technical-report.tex
work in progress - trying to figure out what we wanted with return pointers. At the moment I am trying to introduce a code and data return pointer (I think that is what we ended up with).

@@
 \newcommand{\rret}{\rO}
+\newcommand{\rretc}{\mathrm{r}_\mathrm{ret code}}
+\newcommand{\rretd}{\mathrm{r}_\mathrm{ret data}}
 \newcommand{\rdata}{\mathrm{r}_\mathrm{data}}
@@
     &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
-    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\     &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptr}\\ 
-    &\RegName & \defbnf &\pcreg \mid \rO \mid \rstk \mid \rdata \mid \dots \\
+    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
+    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
+    &\RegName & \defbnf &\pcreg \mid \rretc \mid \rretd \mid \rstk \mid \rdata \mid \dots \\
     &\Reg & \defeq & \RegName \fun \Word\\
@@
     &\MemSeg & \defeq & \Addr \parfun \Word \\
-    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
-    {\sourcecolor{} \sstk \in}& {\sourcecolor{} \Stack} & {\sourcecolor{} \defeq} & {\sourcecolor{} \StkFrame^*} \\
+    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \Caps \times \MemSeg \times \Addr}\\
+    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
     \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
@@
 \item $\cbnf$ defines domain $\Caps$
+\item $\Perm$ is defined as the permissions in Figure~\ref{fig:perm-hier}.
 \end{itemize}
+
+In the source language, $\src{\Stack}$ is a call stack that contains the data for each call.
+The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) the old data capability, 3) caller private stack, and 4) the address of the old stack pointer.
 
@@
     n & \in & \nats \\
-    \src{r} &\in &  \RegName \setminus \{\rO,\rstk\}\\
+    \src{r} &\in &  \RegName \setminus \{\rretc, \rretd, \rstk\}\\
     \src{\rn} &\defbnf & \src{r} \mid n \\
@@
 \[
-  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rO, \rdata \} \ldotp \reg(r) \neq \stkptr(\_)
+  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rretc, \rretd, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
 \]
@@
 \[
-  \noRetStkMs{\ms} \defeq \defeq \forall \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) \neq \stkptr(\_)
+  \noRetStkMs{\ms} \defeq \forall \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) \neq \stkptr{\_}
 \]
@@
                                         \begin{array}[t]{l}
-                                        \update{\pcreg}{\cb_1}\update{\rstk}{\stkptr{\perm,[\aaddr,eaddr],\aaddr}}\\
-                                        \update{\rO}{\retptr}\update{\rdata}{\cb_2} \text{ and }
+                                        \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
+                                        \update{\rretc}{\retptrc}\\
+                                        \update{\rretd}{\retptrd}\\
+                                        \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
                                         \end{array}
@@
                                         \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
-                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
+                                        \stk' = (\opc,\Phi(\rdata),\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
                                         \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
@@
 \end{align*}
+\lau{do we need to require anything on the data capability we put on the call stack?}
 

### 2017-08-31 fully-abstract-comp/technical-report.tex
everything so far. lots of holes here and there.

@@
 \newcommand{\ssem}[2][\Phi]{\sem{#2}_{\mathrm{src}}(#1)}
+\newcommand{\tsem}[2][\Phi]{\sem{#2}_{\mathrm{trg}}(#1)}
 \newcommand{\dom}{\mathrm{dom}}
@@
 \newcommand{\parfun}{\rightharpoonup}
+
+% Text
+\newcommand{\tand}{\text{ and }}
+\newcommand{\totherwise}{\text{otherwise }}
 
@@
 
+\newcommand{\fourinstr}[5]{
+  \ifthenelse{\equal{#2#3#4#5}{}}
+  {\zinstr{#1}}
+  {\zinstr{#1} \; #2 \; #3 \; #4 \; #5}
+}
+
 %%% Source language
@@
 \newcommand{\sjmp}[1]{\oneinstr{\src{jmp}}{#1}}
-\newcommand{\ssetatob}[1]{\oneinstr{\src{seta2b}}{#1}}
 \newcommand{\spush}[1]{\oneinstr{\src{push}}{#1}}
@@
 \newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
+\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
 
@@
 \newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
-\newcommand{\ssubseg}[3]{\threeinstr{\src{subseg}}{#1}{#2}{#3}}
 \newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
@@
 \newcommand{\scseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}
+
+% Four arguments
+\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
 
@@
 \newcommand{\tjmp}[1]{\oneinstr{\trg{jmp}}{#1}}
-\newcommand{\tsetatob}[1]{\oneinstr{\trg{seta2b}}{#1}}
 
@@
 \newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
+\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
 
@@
 \newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
-\newcommand{\tsubseg}[3]{\threeinstr{\trg{subseg}}{#1}{#2}{#3}}
 \newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
@@
 \newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}
+
+% Four arguments
+\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
 
@@
 \newcommand{\opc}{\var{opc}}
+\newcommand{\odata}{\var{odata}}
 \newcommand{\vsc}{\var{sc}}
@@
 }
+\newcommand{\encPerm}[1]{\plainfun{encPerm}{#1}}
 \newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
@@
 \newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
+\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
+\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
+\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
+\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
 \newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
+\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
 \newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
@@
     &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
-    &\RegName & \defbnf &\pcreg \mid \rretc \mid \rretd \mid \rstk \mid \rdata \mid \dots \\
+    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
     &\Reg & \defeq & \RegName \fun \Word\\
@@
     &\MemSeg & \defeq & \Addr \parfun \Word \\
-    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \Caps \times \MemSeg \times \Addr}\\
+    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
     \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
@@
 In the source language, $\src{\Stack}$ is a call stack that contains the data for each call.
-The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) the old data capability, 3) caller private stack, and 4) the address of the old stack pointer.
+The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) caller private stack, and 3) the address of the old stack pointer.
 
@@
 \]
-where $r\in \RegName$
+where $r\in \RegName$.
 
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
-      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
+      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
 \end{array}
@@
     n & \in & \nats \\
-    \src{r} &\in &  \RegName \setminus \{\rretc, \rretd, \rstk\}\\
+    \src{r} &\in &  \RegName \setminus \{\rret, \rstk\}\\
     \src{\rn} &\defbnf & \src{r} \mid n \\
@@
 
-\subsubsection{Helpful functions}
+\subsubsection{Helpful functions and sets}
+
 \[
@@
 \[
+  \readAllowed{} \defeq \dots
+\]
+
+\[
+  \writeAllowed{} \defeq \dots
+\]
+
+\[
+  \writeLocalAllowed{} \defeq \dots
+\]
+
+\[
+  % Lau: this should work for everything that might be local - used in store.
+  \isLoc{c} \defeq \dots
+\]
+
+\[
+  \nonLoc{c} \defeq \dots
+\]
+
+\[
   \nonExec{\cb} \defeq \dots
@@
 \[
-  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rretc, \rretd, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
+  \opaquePerm{} \defeq \dots
+\]
+
+\[
+  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rret, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
 \]
@@
                                       \begin{array}[t]{l}
-                                        % "red conditions" (reference to blackboard picture)
+                                        % "red conditions" (in reference to blackboard picture)
                                         \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
@@
                                         \nonLoc{\cb_1,\cb_2} \text{ and } \\
-                                        \reg' = \reg
+                                        \reg' = \reg''
                                         \begin{array}[t]{l}
                                         \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
-                                        \update{\rretc}{\retptrc}\\
-                                        \update{\rretd}{\retptrd}\\
+                                        \update{\rretc}{\retptr}\\
                                         \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
@@
 \\
+                                        \text{where $\forall r \in \RegName \ldotp \nonLoc{\reg''(r)} \text{ and } \reg''(r) = \reg(r)$}\\
+                                        \quad \text{$\text{ or } \reg''(r)=0 $}\\
                                         \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
                                         \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
-                                        \stk' = (\opc,\Phi(\rdata),\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
+                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
                                         \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
@@
 \end{align*}
-\lau{do we need to require anything on the data capability we put on the call stack?}
 
 The $\scseal{}{}{}$ instruction seals a sealable capability with a seal.
+The base of the seal is used for the seal.
+The used seal only works if it still grants authority to seal (that is the base of the seal is smaller than or equal to the end).
 \begin{align*}
@@
                                   \nonExec{c_2}
-                                \end{array}\\
+                                \end{array}\\\\
 %
-                                \\
+                                (\mem,\reg',\stk',\ms_\stk') &
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_1) = \retptrc \text{ and } \\
+                                  \Phi(r_2) = \retptrd \text{ and } \\
+                                  \Phi = (\mem, \reg, \stk, \ms_\stk) \text{ and } \\
+                                  \stk = (\opc,\ms_\priv,\aaddr_\stk) :: \stk' \text{ and } \\
+                                  \reg \arraycolsep=0pt
+                                  \begin{array}[t]{l}
+                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\\
+                                  \end{array}
+                                \end{array}\\\\
 %
@@
 \end{align*}
-\lau{This is $\sxjmp{}{}$ for the target language - this one also needs to allow jumps to the return pointer}
+\lau{questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}
+
 Notice, $\rdata$ serves the same function as the IDC register in CHERI, i.e.\ this is where the data capability is put.
@@
 
+\begin{align*}
+  \ssem{\sload{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \dots & \\
+                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}} &
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
+                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                  \perm \in \readAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand\\
+                                  \var{w} = \ms_\stk(\addr) 
+                                \end{array}\\
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \ssem{\sstore{r_1}{r_2}} = & 
+                               \begin{cases}
+                                 \dots & \\
+                                 \updPcAddr{\Phi\update{\ms_\stk.a}{w}} & 
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi = (\_, \_, \_, \ms_\stk) \tand \\
+                                   \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                   \perm \in \writeAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand \\
+                                   \aaddr \in \dom(\ms_\stk) \tand \\
+                                   \var{w} = \Phi(r_2) \tand \\
+                                   \text{if } \isLoc{\var{w}} \text{,} \\
+                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                 \end{array}
+                               \end{cases}
+\end{align*}
+
 \begin{itemize}
@@
 \begin{align*}
-  \ssem{\sxjmp{r_1}{r_2}} = & 
+  \tsem{\tfail} = & \failed
+\end{align*}
+
+\begin{align*}
+  \tsem{\thalt} = & \halted
+\end{align*}
+
+\begin{align*}
+  \tsem{\tjmp{r}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tjnz{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tisptr{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgeta{r_1}{r_2}} = &
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{a}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_),\_,\_,a) \text{ and }\\
+                                  \perm \not\in \opaquePerm{}
+                                \end{array} \\
+                                \failed & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgetb{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgete{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgetp{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_1) = ((\perm,\_),\_,\_,\_)\\
+                                \failed & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgetl{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tmove{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tstore{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tload{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tlea{r_1}{r_2}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\trestrict{r_1}{r_2}{r_3}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tsubseg{r_1}{r_2}{r_3}{r_4}} = &
+                                     \begin{cases}
+                                       \updPcAddr{\Phi\update{r_1}{\var{c}}} &
+                                       \arraycolsep=0pt
+                                       \begin{array}[t]{l}
+                                         \Phi(r_2) = ((\perm,\gl),\baddr, \eaddr, \aaddr) \tand \\
+                                         \text{for $i \in \{3,4\}$ and} \\
+                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
+                                         \quad \text{in either case $n_i \in \ints$ and} \\
+                                         \quad \baddr \leq n_3 \tand \\
+                                         \quad n_4 \leq \eaddr \text{ or } n_4 = \infty \tand \\
+                                         \quad \perm \neq \enter \tand \\
+                                         \quad c = ((\perm,\gl),n_3,n_4,\aaddr)
+                                       \end{array} \\
+                                       \updPcAddr{\Phi\update{r_1}{\var{s}}} &
+                                       \arraycolsep=0pt
+                                       \begin{array}[t]{l}
+                                         \Phi(r_2) = \seal{\perm,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
+                                         \text{for $i \in \{3,4\}$ and} \\
+                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
+                                         \quad \text{in either case $n_i \in \ints$ and} \\
+                                         \quad \phi_{\var{base}} \leq n_3 \tand \\
+                                         \quad n_4 \leq \phi_{\var{end}} \text{ or } n_4 = \infty \tand \\
+                                         \quad \perm \neq \enter \tand \\
+                                         \quad c = \seal{\perm, n_3,n_4}
+                                       \end{array} \\
+                                       \failed & \totherwise
+                                     \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tlt{r_1}{r_2}{r_3}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tplus{r_1}{r_2}{r_3}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tminus{r_1}{r_2}{r_3}} = & \dots
+\end{align*}
+
+\begin{align*}
+  \tsem{\tsetatob{r_1}{r_2}} = & 
+                                 \begin{cases}
+                                   \updPcAddr{\Phi\update{\reg.r_1}{c}} &
+                                   \arraycolsep=0pt
+                                   \begin{array}[t]{l}
+                                     \Phi(r_2) = ((\perm,\gl),\baddr,\eaddr,\aaddr) \tand \\
+                                     c = ((\perm,\gl),\baddr,\eaddr,\baddr) \\
+                                     \perm \neq \enter \tand \perm \not\in \opaquePerm{}
+                                   \end{array}
+                                   \\
+                                   \failed & \totherwise
+                                 \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\txjmp{r_1}{r_2}} = & 
                               \begin{cases}
@@
                                 \begin{array}[t]{l}
-                                  [\reg.\pcreg \mapsto c_1] \\
-                                  [\reg.\rdata \mapsto c_2]
+                                  \update{\reg.\pcreg}{c_1} \\
+                                  \update{\reg.\rdata}{c_2}
                                 \end{array} & 
@@
 
+Sealing a capability: $\tcseal{}{}{}$ seals a sealable capability
+\begin{align*}
+  \tsem{\tcseal{r_1}{r_2}{r_3}} = & 
+                                    \begin{cases}
+                                      \updPcAddr{\Phi\update{\reg.r_1}{\sealed{\Phi(r_2),\phi_{\var{base}}}}} &
+                                      \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        \Phi(r_2) \in \SealableCaps \text{ and } \\
+                                        \Phi(r_3) = \seal{\gl,\phi_{\var{base}},\phi_{\var{end}}} \\
+                                        \phi_{\var{base}} \leq \phi_{\var{end}} % This allows one to restrict a seal to not have any authority
+                                      \end{array} \\
+                                      \failed & \text{otherwise}
+                                    \end{cases}
+\end{align*}
+
+\subsection{Target program linking and layout}
+For a $\scall{}{}$ to be secure after compilation, we need to use the seal mechanism.
+In order for a program to seal a capability, it must have a seal available.
+We propose that a compiled program must have the layout sketched in Figure~\ref{fig:trg-prog-link}.
+The first address of a program contains a seal (unique to this program), the second address contains a capability for a linking table, and the following addresses contain the instructions of the program.
+The capability for the program points to the first address of the program and it has permission no greater then global read-execute (notice that it does not have write).
+
+The capability for the linking table is a global read-write capability, which means that it does not grant access to alter the table (which means that the table could potentially be shared between multiple parties).
+The linking table contains pairs of sealed code and data capabilities.
+
+The program should also have access to a persistent memory which it has in the register $\rdata$ where it has a global read-write capability (or something less) (notice that this capabilty should not have execute permission).
+Finally the program also has access to local memory in the form of a stack.
+The capability for the stack resides in $\rstk$ and the capability should have local read-write-local permission (or less)
+\lau{The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}
+
+\lau{should the capability for the ``persistent'' memory reside along side the code?}
+
+Notice that the linking part is not strictly necessary for the compilation, but we would like to be able to link our known programs with other programs.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{img/prog_layout.jpg}
+  \caption{Target program linking and layout (there is a typo in the figure: the pc capability should be global rather than local).}
+  \label{fig:trg-prog-link}
+\end{figure}
+
 \subsection{Compiler}
@@
 \]
+
+\[
+  \comp{\scall{r_1}{r_2}} = 
+  \begin{array}[t]{l}
+    % Make the stack capability into the return capability and seal it:
+    %   Push r_data to the stack (should this be made the responsibility of the programmer?
+If it happens here, then it should probably also be part of the call semantics.
+Especially, the call semantics should "eat" another piece of memory on the stack to make sure there is no discrepancy.)
+    %   Seal the stack capability (with what?
+where does the seal come from?
+I think we talked about this, and it is something that should be handled like linking (a fresh seal is available in the beginning of a programs code along with a capability for a linking table)).
+    % restrict the stack capability to the unused part (and clear this part).
+    % make return capability and ceal it:
+    %   Move a copy of pc to a different register.
+    %   Adjust the pc copy to point to the return address.
+    %   Seal it with the seal.
+    % remove all data from temp registers (that is at least the seal).
+    % jmpx to r1 and r2
+    % restore code:
+    %   move the stack pointer to the proper register
+    %   pop the data capability
+  \end{array}
+\]
+
 \clearpage

### 2017-09-01 fully-abstract-comp/technical-report.tex
Fix LaTeX compilation.
added linear capabilities and split/splice instructions w/ semantics

@@
 \newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
+\newcommand{\ssplice}[2]{\twoinstr{\src{splice}}{#1}{#2}}
+\newcommand{\ssplit}[2]{\twoinstr{\src{split}}{#1}{#2}}
 
@@
 \newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
-\newcommand{\scseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}
+\newcommand{\scseal}[3]{\threeinstr{\src{cseal}}{#1}{#2}{#3}}
 
@@
 \newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
+\newcommand{\tsplice}[2]{\twoinstr{\trg{splice}}{#1}{#2}}
+\newcommand{\tsplit}[2]{\twoinstr{\trg{split}}{#1}{#2}}
 
@@
 \newcommand{\Global}{\shareddom{Global}}
+\newcommand{\Linear}{\shareddom{Linear}}
 \newcommand{\MemSeg}{\shareddom{MemorySegment}}
@@
 \newcommand{\glbnf}{\shareddom{gl}}
+\newcommand{\linbnf}{\shareddom{l}}
 \newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
@@
 \newcommand{\gl}{\shareddom{g}}
+\newcommand{\lin}{\shareddom{l}}
 \newcommand{\base}{\shareddom{base}}
@@
 
+%%% linearity
+\newcommand{\plainlinearity}[1]{\mathrm{#1}}
+\newcommand{\linear}{\plainlinearity{linear}}
+\newcommand{\normal}{\plainlinearity{normal}}
+
 %%% Permissions
@@
     &\glbnf & \defbnf & \glob \mid \local \\
-    & \aendbnf & \in & \Addr \uplus \{\infty \} \\
-    &  \sealendbnf & \in & \Seal \uplus \{\infty \} \\
-    &\scbnf &\defbnf & ((\permbnf,\glbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
+    &\linbnf & \defbnf & \linear \mid \normal \\
+    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
+    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
+    &\scbnf &\defbnf & ((\permbnf,\glbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
@@
 \item $\glbnf$ defines domain $\Global$
+\item $\linbnf$ defines domain $\Linear$
 \item $\scbnf$ defines domain $\SealableCaps$
@@
       & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
-      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{r}{r} \mid \tcseal{r}{r}{r}
+      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
+      & & \tsplice{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}
 \end{array}
@@
 \end{figure}
+\lau{We should maybe add a picture with local/global hieararchy.}
+\lau{For now, I will add the linear capabilities such that a linear capability stays linear.
+It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely).
+One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}
 
@@
 \item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
+\item Linear capabilities are cleared when they move around in memory.
 \end{itemize}
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\_,\_,a) \text{ and }\\
+                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,a) \text{ and }\\
                                   \perm \not\in \opaquePerm{}
@@
                               \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_1) = ((\perm,\_),\_,\_,\_)\\
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_1) = ((\perm,\_,\_),\_,\_,\_)\\
                                 \failed & \totherwise
@@
                                        \begin{array}[t]{l}
-                                         \Phi(r_2) = ((\perm,\gl),\baddr, \eaddr, \aaddr) \tand \\
+                                         \Phi(r_2) = ((\perm,\gl,\lin),\baddr, \eaddr, \aaddr) \tand \\
                                          \text{for $i \in \{3,4\}$ and} \\
@@
                                          \quad \perm \neq \enter \tand \\
-                                         \quad c = ((\perm,\gl),n_3,n_4,\aaddr)
+                                         \quad c = ((\perm,\gl,\lin),n_3,n_4,\aaddr)
                                        \end{array} \\
@@
                                    \begin{array}[t]{l}
-                                     \Phi(r_2) = ((\perm,\gl),\baddr,\eaddr,\aaddr) \tand \\
-                                     c = ((\perm,\gl),\baddr,\eaddr,\baddr) \\
+                                     \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                     c = ((\perm,\gl,\lin),\baddr,\eaddr,\baddr) \\
                                      \perm \neq \enter \tand \perm \not\in \opaquePerm{}
@@
                                       \failed & \text{otherwise}
+                                    \end{cases}
+\end{align*}
+
+$\tsplice{}{}$ we allow two capabilities to be spliced if the second one starts on the address after the first one ends.
+They must also agree on permission, globality and linearity (so we also allow non-linear capabilities to be spliced).
+We do, however, exclude enter capabilities as this may cause subtle security issues.
+\begin{align*}
+  \tsem{\tsplice{r_1}{r_2}} = &
+                                \begin{cases}
+                                  \updPcAddr{\Phi\update{r_1}{\dots}\update{r_2}{c}} &
+                                  \arraycolsep=0pt
+                                  \begin{array}[t]{l}
+                                    \Phi(r_1) = ((\perm, \gl, \lin), \baddr_1, \eaddr_1, \aaddr_1) \\
+                                    \Phi(r_2) = ((\perm, \gl, \lin), \baddr_2, \eaddr_2, \aaddr_2) \\
+                                    \eaddr_1 + 1 = \baddr_2 \\
+                                    \perm \neq \enter \\
+                                    c = ((\perm, \gl, \lin), \baddr_1, eaddr_2, \aaddr_1)
+                                  \end{array}
+                                  \\
+                                  \failed & \totherwise
+                                \end{cases}
+\end{align*}
+
+$\tsplit{}{}$ this instruction splits a capability in two.
+We require the address we wish to split on to be in the interval of the authority granted by the capability to prevent handing out authority from capabilities with no authority.
+We prohibit enter capabilities to prevent subtle errors.
+\begin{align*}
+  \tsem{\tsplit{r_1}{r_2}} = &
+                               \begin{cases}
+                                 \updPcAddr{\Phi\update{r_1}{c_1}\update{r_2}{c_2}} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi(r_1) = ((\perm, \gl, \lin), \baddr, eaddr, \aaddr) \\
+                                   r_2 = n \text{ or } \Phi(r_2) = n \text{ and in either case } n \in \nats\\
+                                   \perm \neq \enter\\
+                                   \baddr \leq n  < \eaddr\\
+                                   c_1 = ((\perm,\gl,\lin),\baddr, n, \aaddr) \\
+                                   c_2 = ((\perm,\gl,\lin),n + 1, \eaddr,\aaddr)
+                                 \end{array} \\
+                                 \failed & \totherwise
                                \end{cases}

### 2017-09-04 fully-abstract-comp/technical-report.tex
updated some of the semantics (load and store) to take linear capabilities into consideration.

@@
 \newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
+\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
 \newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
+\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
 \newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
 \newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
+\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
 
@@
     \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
-    &\Word & \defeq & \ints \uplus \Caps\\
+    &\Word & \defeq & \ints \uplus \SealableCaps\\
     \permbnf \in& \Perm & \defbnf & \dots \\
@@
 \item $\cbnf$ defines domain $\Caps$
-\item $\Perm$ is defined as the permissions in Figure~\ref{fig:perm-hier}.
+\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
 \end{itemize}
@@
 In the source language, $\src{\Stack}$ is a call stack that contains the data for each call.
-The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) caller private stack, and 3) the address of the old stack pointer.
+The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) caller's private stack, and 3) the address of the old stack pointer.
 
@@
 \subsection{Syntax}
-The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (Inspired by CHERI).
+The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI).
 The syntax of the instructions of the target machine is defined as follows:
@@
     n & \in & \nats \\
-    \src{r} &\in &  \RegName \setminus \{\rret, \rstk\}\\
+    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
     \src{\rn} &\defbnf & \src{r} \mid n \\
@@
 \]
+\lau{We had $\rret$ as not being one of the registers on the source machine.
+I could not recall why we had this, so I removed it.
+(I think it may have been from when we had call and return instructions?)}
 There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
@@
 \[
-  \updPcAddr{} \defeq \dots
-\]
-
-\[
-  \updPcPerm{} \defeq \dots
+  \updPcAddr{c} \defeq \dots
+\]
+
+\[
+  \updPcPerm{c} \defeq \dots
 \]
@@
 \[
+  \isLinear{\cb} \defeq \dots
+\]
+
+\[
+  \nonLinear{\cb} \defeq \dots
+\]
+
+\[
   \nonLoc{c} \defeq \dots
@@
   \noRetStkMs{\ms} \defeq \forall \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) \neq \stkptr{\_}
+\]
+
+\[
+  \withinBounds{c} \defeq \dots
 \]
@@
 
+\lau{how should the interaction between read-only memory and linear capabilities be?
+That is, if a linear capability resides in read-only memory and you attempt to load it what happens?
+Do the load succeed with the memory cell being cleared or do the load fail (because we need to clear the cell, but we do not have the necessary permissions)?
+}
 \begin{align*}
@@
                               \begin{cases}
-                                \dots & \\
+                                \updPcAddr{\update{\reg.r_1}{\var{w}}} &
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi = (\mem, \_, \_, \_) \tand \\
+                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                  \perm \in \readAllowed{} \tand \withinBounds{\var{c}} \tand \\
+                                  \var{w} = \mem(\aaddr) \tand \\
+                                  \nonLinear{\var{w}}
+                                \end{array}\\
+                                \updPcAddr{\update{\reg.r_1}{\var{w}}}\update{\mem(\aaddr)}{0} &
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi = (\mem, \_, \_, \_) \tand \\
+                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                  \perm \in \readAllowed{} \cup \writeAllowed{} \\
+                                  \tand \withinBounds{\var{c}} \tand \\
+                                  \var{w} = \mem(\aaddr) \tand \\
+                                  \isLinear{\var{w}}
+                                \end{array}\\
+                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}\update{\ms_\stk(\aaddr)}{0}} & 
+                                % linear capability
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \begin{array}[t]{l}
+                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
+                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                  \perm \in \readAllowed{} \cup \writeAllowed{} \tand \\
+                                  \baddr \leq \aaddr \leq \eaddr \tand \var{w} = \ms_\stk(\aaddr) \tand \\
+                                  \isLinear{\var{w}}
+                                \end{array}
+                                \\
+                                \end{array}
+                                \\
                                 \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}} &
                                 \arraycolsep=0pt
+                                % non-linear capability
                                 \begin{array}[t]{l}
@@
                                   \perm \in \readAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand\\
-                                  \var{w} = \ms_\stk(\addr) 
+                                  \var{w} = \ms_\stk(\aaddr) \tand \\
+                                  \nonLinear{\var{w}}
                                 \end{array}\\
+                                \failed & \totherwise
                               \end{cases}
@@
                                \begin{cases}
-                                 \dots & \\
-                                 \updPcAddr{\Phi\update{\ms_\stk.a}{w}} & 
+                                 \updPcAddr{\Phi\update{\ms.a}{w}}\update{\reg.r_2}{0} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi(r_1) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                   \perm \in \writeAllowed{} \tand \withinBounds{\var{c}} \tand \\
+                                   \var{w} = \Phi(r_2) \tand \\
+                                   \isLinear{\var{w}} \tand \\
+                                   \text{if } \var{w} = ((\_,\local,\_),\_,\_,\_) \\
+                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                 \end{array} \\
+                                 \updPcAddr{\Phi\update{\ms.a}{w}} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi(r_1) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                   \perm \in \writeAllowed{} \tand \withinBounds{\var{c}} \tand \\
+                                   \var{w} = \Phi(r_2) \tand \\
+                                   \nonLinear{\var{w}} \tand \\
+                                   \text{if } \var{w} = ((\_,\local,\_),\_,\_,\_) \\
+                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                 \end{array} \\
+%
+                                 \updPcAddr{\Phi\update{\ms_\stk.a}{w}}\update{reg.r_2}{0} & 
                                  \arraycolsep=0pt
@@
                                    \var{w} = \Phi(r_2) \tand \\
+                                   \isLinear{w} \tand \\
                                    \text{if } \isLoc{\var{w}} \text{,} \\
                                    \text{ then } \perm \in \writeLocalAllowed{}
-                                 \end{array}
+                                 \end{array} \\
+                                 \updPcAddr{\Phi\update{\ms_\stk.a}{w}} & 
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi = (\_, \_, \_, \ms_\stk) \tand \\
+                                   \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                   \perm \in \writeAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand \\
+                                   \aaddr \in \dom(\ms_\stk) \tand \\
+                                   \var{w} = \Phi(r_2) \tand \\
+                                   \nonLinear{w} \tand \\
+                                   \text{if } \isLoc{\var{w}} \text{,} \\
+                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                 \end{array} \\
+                                 \failed & \totherwise
                                \end{cases}
@@
 \begin{align*}
-  \tsem{\tmove{r_1}{r_2}} = & \dots
+  \tsem{\tmove{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
+                                \updPcAddr{\Phi\update{\reg.r_1}{\Phi(r_2)}}\update{\reg.r_2}{0} & \totherwise 
+                              \end{cases}
 \end{align*}

### 2017-09-06 fully-abstract-comp/technical-report.tex
Small note I forgot to commit.
written up some of the semantics for the language

@@
 \newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
-\newcommand{\sisptr}[2]{\twoinstr{\src{isptr}}{#1}{#2}}
+\newcommand{\sisptr}[2]{\twoinstr{\src{gettype}}{#1}{#2}}
 \newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
@@
 \newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
-\newcommand{\sgetl}[2]{\twoinstr{\src{getl}}{#1}{#2}}
+\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
+\newcommand{\sgetlin}[2]{\twoinstr{\src{getlin}}{#1}{#2}}
 \newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
@@
 \newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
-\newcommand{\tisptr}[2]{\twoinstr{\trg{isptr}}{#1}{#2}}
+\newcommand{\tisptr}[2]{\twoinstr{\trg{gettype}}{#1}{#2}}
 \newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
@@
 \newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
-\newcommand{\tgetl}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
+\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
+\newcommand{\tgetlin}[2]{\twoinstr{\trg{getlin}}{#1}{#2}}
 \newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
@@
 \newcommand{\update}[2]{[#1 \mapsto #2]}
-\newcommand{\updReg}[3][\Phi]{#1\update{\reg.#2}{#3}}
-\newcommand{\updPc}[2][\Phi]{\updReg[#1]{\pcreg}{#2}}
+\newcommand{\updReg}[2]{\update{\reg.#1}{#2}}
+\newcommand{\updPc}[1]{\Phi\updReg{\pcreg}{#1}}
 
@@
 \newcommand{\sastk}{\shareddom{a_{stk}}}
-\newcommand{\perm}{\shareddom{perm}}
-\newcommand{\gl}{\shareddom{g}}
-\newcommand{\lin}{\shareddom{l}}
+\newcommand{\perm}{\var{perm}}
+\newcommand{\gl}{\var{g}}
+\newcommand{\lin}{\var{l}}
 \newcommand{\base}{\shareddom{base}}
@@
 \newcommand{\failed}{\mathrm{failed}}
-\newcommand{\undefined}{\mathrm{undefined}}
+% DOMI: defining a macro named ``undefined'' breaks many latex packages.
+%       (this is just another way that LaTeX is broken as a programming language)
+% \newcommand{\undefined}{\mathrm{undefined}}
+\newcommand{\tundefined}{\mathrm{undefined}}
 \newcommand{\halted}{\mathrm{halted}}
@@
 }
-\newcommand{\encPerm}[1]{\plainfun{encPerm}{#1}}
+
+\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
+\newcommand{\encPerm}[1]{\plainfun{encocePermission}{#1}}
 \newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
@@
 \newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
+\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
 
@@
 \]
+\lau{Should seals also have the option to be linear?}
 The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
@@
 \trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
-      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetl{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
+      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetloc{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
       & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
@@
 
+\[
+  \nonZero{c} \defeq \dots
+\]
+
 \subsubsection{Source Language}
-\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update.}
+\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated. It was $\sxjmp{}{}$ that needed an update.}
+
 \textbf{Call}\\
@@
                                         \text{where $\forall r \in \RegName \ldotp \nonLoc{\reg''(r)} \text{ and } \reg''(r) = \reg(r)$}\\
-                                        \quad \text{$\text{ or } \reg''(r)=0 $}\\
+                                        \quad \text{$\text{ or } \reg''(r)=0 $} \tand \\
+                                        \text{ for $i \in \{1,2\}$ if $\isLinear{\Phi(r_i)}$, then $\reg''(r_i) = 0$ and} \\
                                         \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
@@
                                       & \\
-      \undefined & \arraycolsep=0pt
+      \tundefined & \arraycolsep=0pt
                                       \begin{array}[t]{l}
@@
                                     \begin{cases}
-                                      \updReg{r_1}{\vsc} & 
+                                      \Phi\updReg{r_1}{\vsc} & 
                                       \arraycolsep=0pt
@@
                               \begin{cases}
-                                \Phi
+                                \Phi'
                                 \arraycolsep=0pt
@@
                                   \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
+                                  \text{ for $i \in \{1,2\}$ if $\isLinear{c_i}$, then $\Phi'(r_i) = 0$} \\
+                                  \text{ otherwise $\Phi'(r_i) = \Phi(r_i)$ and} \\
+                                  \text{ $ \Phi(r) = \Phi'(r) \text{ for } r\in \tRegName \setminus \{r_1,r_2\}$ and} \\
                                   \nonExec{c_2}
@@
                                   \begin{array}[t]{l}
-                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\\
+                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\update{r_2}{0}\\
                                   \end{array}
@@
 
-\begin{align*}
-  \ssem{\sjmp{r}} = & \dots
+\textbf{jmp}\\
+\begin{align*}
+  \ssem{\sjmp{r}} = & 
 \end{align*}
@@
                                 \end{array}\\
-                                \updPcAddr{\update{\reg.r_1}{\var{w}}}\update{\mem(\aaddr)}{0} &
+                                \updPcAddr{\update{\reg.r_1}{\var{w}}\update{\mem(\aaddr)}{0}} &
                                 \arraycolsep=0pt
@@
 
+\textbf{Getters: geta, getb, gete, getp, getloc, and getlin}
+
+\begin{align*}
+  \tsem{\tgeta{r_1}{r_2}} = &
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{\aaddr}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,\aaddr) \text{ or } \\
+                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} \tand\\
+                                  \perm \not\in \opaquePerm{}
+                                \end{array} \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\sgetb{r_1}{r_2}} = &
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{\baddr}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_,\_),\baddr,\_,\_) \text{ or } \\
+                                  \quad \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} \tand\\
+                                  \perm \not\in \opaquePerm{}
+                                \end{array} \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\sgete{r_1}{r_2}} = &
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{\eaddr}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_,\_),\_,\eaddr,\_) \text{ or } \\
+                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} \tand\\
+                                  \perm \not\in \opaquePerm{}
+                                \end{array} \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\lau{Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}
+
+\begin{align*}
+  \tsem{\sgetp{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_,\_),\_,\_,\_) \text{ or } \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}\\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\sgetloc{r_1}{r_2}} = &                              
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\gl}) & \Phi(r_2) = ((\_,\gl,\_),\_,\_,\_) \\
+                                \updPcAddr{\Phi\updReg{r_1}{\local}} & \Phi(r_2) = \stkptr{\_,\_,\_,\_}\\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\sgetlin{r_1}{r_2}} = &                              
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\_,\lin),\_,\_,\_) \\
+                                \updPcAddr{\Phi\updReg{r_1}{\linear}} & \Phi(r_2) = \stkptr{\_,\_,\_,\_}\\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\textbf{gettype}\\
+\lau{I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
+We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
+\lau{TODO: Write out what $\encType{}$ should satisfy}
+\begin{align*}
+  \tsem{\sisptr{r_1}{r_2}} = & 
+                               \begin{cases}
+                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
+                               \end{cases}
+\end{align*}
+
 \begin{itemize}
@@
 
-\begin{align*}
-  \tsem{\tjmp{r}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tjnz{r_1}{r_2}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tisptr{r_1}{r_2}} = & \dots
-\end{align*}
-
+\textbf{jmp and jnz}\\
+Notice that the order of the updates of the configuration actually matters here.
+As it is below, a successful jump to $\pcreg$ results in a loop, but if the order of the updates to the configuration had been reversed, then the execution would fail as the $\pcreg$ register would contain 0.
+\begin{align*}
+  \tsem{\tjmp{r}} = & 
+                      \begin{cases}
+                        \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_r)}
+                             & \nonLinear{\Phi(r)} \\
+                        \Phi\updReg{\pcreg}{\Phi(r)} & \totherwise
+                      \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tjnz{r_1}{r_2}} = &       
+                             \begin{cases}
+                               \Phi\updReg{\pcreg}{r_1} & \text{if $\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
+                               \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_1)}
+                                               & \arraycolsep=0pt                               
+                                               \begin{array}[t]{l}
+                                                 \isLinear{\Phi(r)} \tand
+                                                 \nonZero{\Phi{r_2}}
+                                               \end{array}
+\\
+                               \updPcAddr{\Phi} & \totherwise
+                             \end{cases}
+\end{align*}
+
+\textbf{Getters: geta, getb, gete, getp, getloc, and getlin}
+Notice the difference between the getters in the target languages and their equivalents in the source language is that these do not have a case for the stack pointer as it does not exist in the language.
 \begin{align*}
@@
                               \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{a}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,a) \text{ and }\\
+                                \updPcAddr{\Phi\update{\reg.r_1}{\aaddr}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,\aaddr) \tand \\
                                   \perm \not\in \opaquePerm{}
                                 \end{array} \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgetb{r_1}{r_2}} = &
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{\baddr}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_,\_),\baddr,\_,\_) \tand \\
+                                  \perm \not\in \opaquePerm{}
+                                \end{array} \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgete{r_1}{r_2}} = &
+                              \begin{cases}
+                                \updPcAddr{\Phi\update{\reg.r_1}{\eaddr}} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\_,\_),\_,\eaddr,\_) \tand \\
+                                  \perm \not\in \opaquePerm{}
+                                \end{array} \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\lau{Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}
+
+\begin{align*}
+  \tsem{\tgetp{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_,\_),\_,\_,\_)\\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgetloc{r_1}{r_2}} = &                              
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\gl}) & \Phi(r_2) = ((\_,\gl,\_),\_,\_,\_) \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tgetlin{r_1}{r_2}} = &                              
+                              \begin{cases}
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\_,\lin),\_,\_,\_) \\
+                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                              \end{cases}
+\end{align*}
+
+\textbf{gettype}\\
+\lau{I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
+We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
+\lau{TODO: Write out what $\encType{}$ should satisfy}
+\begin{align*}
+  \tsem{\tisptr{r_1}{r_2}} = & 
+                               \begin{cases}
+                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
+                               \end{cases}
+\end{align*}
+
+\textbf{move}\\
+\lau{Like previously, the order of updates in the second case matters for the case where $r_1 = r_2$.}
+\begin{align*}
+  \tsem{\tmove{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
+                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
+                              \end{cases}
+\end{align*}
+
+\begin{align*}
+  \tsem{\tstore{r_1}{r_2}} = & 
+                               \begin{cases}
+                                 \Phi\updReg{r_2}{w_2}
+                                     \update{\mem.a}{\Phi(r_2)} & \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \Phi(r_1) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                       \perm \in \writeAllowed{} \tand \\
+                                       \baddr \leq \aaddr \leq \eaddr \tand \\
+                                       \isLoc{\Phi(r_2)} \Rightarrow \perm \in \writeLocalAllowed{} \tand \\
+                                       \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
+                                       \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2)
+                                     \end{array} \\
                                   \failed & \totherwise
@@
 \begin{align*}
-  \tsem{\tgetb{r_1}{r_2}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tgete{r_1}{r_2}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tgetp{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_1) = ((\perm,\_,\_),\_,\_,\_)\\
-                                \failed & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tgetl{r_1}{r_2}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tmove{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
-                                \updPcAddr{\Phi\update{\reg.r_1}{\Phi(r_2)}}\update{\reg.r_2}{0} & \totherwise 
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tstore{r_1}{r_2}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tload{r_1}{r_2}} = & \dots
+  \tsem{\tload{r_1}{r_2}} = & 
 \end{align*}

### 2017-09-06 fully-abstract-comp/technical-report.tex
Added answer to Dominique. Added approximate timestamps for my comments.

@@
 \]
-\lau{Should seals also have the option to be linear?}
+\lau{06-09-2017: Should seals also have the option to be linear?}
 The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
@@
 \]
-\lau{We had $\rret$ as not being one of the registers on the source machine.
+\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine.
 I could not recall why we had this, so I removed it.
@@
 \end{figure}
-\lau{We should maybe add a picture with local/global hieararchy.}
-\lau{For now, I will add the linear capabilities such that a linear capability stays linear.
+\lau{30-08-2017: We should maybe add a picture with local/global hieararchy.}
+\lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear.
 It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely).
@@
 Futher the code and data capability must be local and there must be no attempt at passing on the stack capability.
-\lau{Add more description here.}
-
-For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}
+\lau{30-08-2017: Add more description here.}
+
+For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{30-08-2017: I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}
 
@@
 \end{align*}
-\lau{questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}
+\lau{30-08-2017: questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}
 
@@
 
-\lau{how should the interaction between read-only memory and linear capabilities be?
+\lau{01-09-2017: how should the interaction between read-only memory and linear capabilities be?
 That is, if a linear capability resides in read-only memory and you attempt to load it what happens?
@@
 
-\lau{Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}
+\lau{01-09-2017: Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}
 
@@
 \textbf{gettype}\\
-\lau{I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
+\lau{06-09-2017: I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
 We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
-\lau{TODO: Write out what $\encType{}$ should satisfy}
+\lau{TODO: 06-09-2017: Write out what $\encType{}$ should satisfy}
 \begin{align*}
@@
 \textbf{move}\\
-\lau{Like previously, the order of updates in the second case matters for the case where $r_1 = r_2$.}
+\lau{06-09-2017: Like previously, the order of updates in the second case matters for the case where $r_1 = r_2$.}
 \begin{align*}
@@
 \end{align*}
+\dominique{31-8-2017: Why do we allow applying subseg to a sealed capability?}
+\lau{6-9-2017: We do not. We allow subseg on the seals that are used to seal capabilities.}
 
@@
 The capability for the stack resides in $\rstk$ and the capability should have local read-write-local permission (or less)
-\lau{The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}
-
-\lau{should the capability for the ``persistent'' memory reside along side the code?}
+\lau{30-08-2017: The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}
+
+\lau{30-08-2017: should the capability for the ``persistent'' memory reside along side the code?}
 

### 2017-09-07 fully-abstract-comp/technical-report.tex
some more semantics (commit before meeting)
Add some notes.

@@
 \subsection{Domains}
+\dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
 Source language domains:
@@
 One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}
+\dominique{7-9-2017: Err...
+you mean that we could allow non-linear capabilities to be made linear, right?
+We definitely do not want to allow the stack or return capability to be made non-linear, as we want to prevent it from being aliased.}
 
@@
 For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{30-08-2017: I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}
+\dominique{7-9-2017: We declare the call to be undefined if the caller is doing something ``stupid'' that would break security but which we do not want to check for.
+This includes: leaving a copy of the stack pointer or the old return pointer in a place accessible to the callee, invoking local callbacks (see discussion in POPL2018 submission of why this is bad).
+The condition that $\cb_2 \neq \stkptr{\_,\_,\_,\_}$ is weird because this would not be the caller's fault, but the callee's.
+Further thought needed regarding correct blame assignment for this case.}
 
@@
 \subsubsection{Target Language}
+\dominique{7-9-2017: It would be nice if we could share definitions for both languages.  Can we factor them in a way that allows this?}
 
@@
 \end{align*}
+\dominique{7-9-2017: add the condition that $r_1 \neq r_2$ for sanity?}
 
@@
   \tsem{\tload{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \Phi\update{\mem.a}{w_2}\updReg{r_1}{w} &
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                  \perm \in \readAllowed{} \tand \\
+                                  \baddr \leq \aaddr \leq \eaddr \tand \\
+                                  w = \Phi.\mem(a) \tand \\
+                                  \isLinear{w} \Rightarrow w_2 = 0 \tand \\
+                                  \nonLinear{w} \Rightarrow w_2 = \Phi.\mem(a)
+                                \end{array} \\
+                                \failed & \totherwise
+                              \end{cases}
 \end{align*}

### 2017-09-07 fully-abstract-comp/technical-report.tex
comments

@@
 \subsection{Domains}
+
 \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
+\lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost).
+I do, however, not think that it is strictly necessary here to make things work.}
 Source language domains:
@@
 We definitely do not want to allow the stack or return capability to be made non-linear, as we want to prevent it from being aliased.}
+\lau{07-09-2017: I actually did mean linear to non-linear (non-linear to linear does not make sense because the non-linear capability may have aliases).
+The operation that makes a linear capability non-linear would have to clear said capability.
+If a linear capability has been made non-linear, then someone who expects a linear capability can observe this difference and behave accordingly, i.e., fail the execution.
+It will, however, remain an invariant that there are no aliases for linear capabilities.}
+\lau{I had forgotten that return capabilities probably need to be linear.
+This would call for a ``load from offset'' operation to be able to load an in code stored seal (see the figure I drew about linking/seals).}
 
@@
 \dominique{7-9-2017: It would be nice if we could share definitions for both languages.  Can we factor them in a way that allows this?}
+\lau{07-09-2017: I agree.
+I will first write out the two languages and then I will look at where they agree (and possibly find some way of unifying the definitions eventhough they don't agree entirely).}
 
@@
 \dominique{7-9-2017: add the condition that $r_1 \neq r_2$ for sanity?}
+\lau{07-09-2017: I would find it odd if the machine had to special case on that (what would it even do? Fail?). As it is now, it follows the invariant that a linear capability is never aliased.}
 

### 2017-09-07 fully-abstract-comp/technical-report.tex
Add some explanation about the observable code size/weird stack instructions.

@@
 \usepackage{array}
+\usepackage{cleveref}
 \usepackage{stmaryrd}
@@
 \subsection{Domains}
+\label{sec:domains}
 
@@
 \subsection{Syntax}
+\label{sec:syntax}
 The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI).
@@
 \subsubsection{Source Language}
+\label{sec:source-language}
 \lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated. It was $\sxjmp{}{}$ that needed an update.}
@@
 \end{enumerate}
+
+\subsection{Some explanation}
+\subsubsection{The source language}
+The whole idea of this paper is to show that we enforce well-bracketed control flow correctly by taking a source language with a native stack and native call and return instructions and well-bracketedness enforced by the operational semantics.
+We then prove that we can compile this source language to a regular assembly target language (without a native stack) in a fully abstract way using our stack and return pointer discipline.
+So the design goals for our source language were precisely these: it should be an intermediate language that stays as close as possible to our target language, can be fully abstractly compiled to the target and feature a native well-behaved stack.
+This source language can then be used as the second-to-last language in the compilation chain of a fully abstract compiler.
+By compiling in a fully abstract way to our source language and then composing with our compiler, fully abstract compilation to our target language follows easily.
+
+These design goals drove us to the current definition of our source language and some special design choices.
+Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
+The reason is that we are working in a language where the length of code blocks can be observable.
+Imagine two trusted blocks of code that looks as follows:
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+p1:       <assert that r1 contains an executable capability to a block of instr_length (push) + instr_length(jmp)>
+          <load return capability to p1-ret1 in r2>
+          <load value 0 in r3>
+          <jump to r1>
+p1-ret1:  <assert that 0 is loaded in the next free stack position>
+          <repeat the above with value 1 in r3>
+          ...
+p1-ret2:  <assert that 1 is loaded in the next free stack position>
+          <repeat the above with value 2 in r3>
+          ...
+p1-ret3:  <assert that 2 is loaded in the next free stack position>
+          halt
+
+p2:       <assert that r1 contains an executable capability to a block of instr_length (push) + instr_length(jmp)>
+          <load return capability to p2-ret1 in r2>
+          <load value 1 in r3>
+          <jump to r1>
+p2-ret1:  <assert that 1 is loaded in the next free stack position>
+          <repeat the above with value 2 in r3>
+          ...
+p2-ret2:  <assert that 2 is loaded in the next free stack position>
+          halt
+\end{lstlisting}
+
+p1 and p2 each accept an (unprotected) closure from the context whose lenght is restricted.
+Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+closure: push r3
+         jmp r2
+\end{lstlisting}
+
+However, this is no longer true in the target language.
+Assume that $\mathtt{push r3}$ is implemented as a store to $\rstk$ followed by an increment of $\rstk$.
+The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+evil-closure: jnz r3 (\pcreg+2)
+              store \rstk r3
+              jmp r2
+\end{lstlisting}
+\dominique{7-9-2017: the above assumes that we have a way of jnz-ing to pc + 2, which may not be true in our actual assembly, but I haven't managed to construct an example that falls entirely within our language.}
+If we now assume that the instructions in evil-closure (jnz+store+jmp) are the same length as the original closure(store+increment+jmp), then the adversary can now construct a closure that discriminates $\mathtt{p1}$ and $\mathtt{p2}$ above.
+The essential point here is that an assumption in the source language that relies on facts about instruction length (``a closure restricted to the instruction-size of push+ret that modifies the stack and returns can only be implemented in one way'') is no longer valid, because there are now more low-level ways of interacting with the stack.
+
+To remedy this, we decided to modify our source language to use a different way of interacting with the stack.
+This alternative approach stays even closer to the target language.
+Specifically, rather than using push/pop and call/return instructions, we do not add any instructions, but instead use special token values that play the role of the stack pointer/return pointer.
+Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the code and data parts of the sealed return closure.
+Rather than invoking push or pop, one would store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:source-language}, this has the semantic effect that one would expect.
+Similarly, rather than invoking return, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.
+Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
+
+Note that this weird behavior is essentially just a weird syntax for something we would otherwise write using call/return and push/pop.
+However, it removes the problem described above of relying on facts that follow from code length, because we can now write exactly the same code in source and target and it will do exactly the same thing if it is well-behaved.
+The code would look exactly the same in source and target, it would just be working with token values in the source versus real stack and return pointers in the target.
+That the problem is really gone should follow from being able to write a full-abstraction proof.
+However, our approach should also simplify that proof.
+The reason is that back-translating should become essentially trivial: rather than performing some form of emulation or a real back-translation (which would inevitably make code size larger), it should be sufficient to just use unmodified target instructions as source instructions that will just be working with different types of values.
+
 \end{document}

### 2017-09-07 fully-abstract-comp/technical-report.tex
added a comment.

@@
 The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+\begin{lstlisting}
+[basicstyle=\sourcecolor{}\ttfamily] 
 p1 ::= if r1 is length 2 then
@@
 Imagine two trusted blocks of code that looks as follows:
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-p1:       <assert that r1 contains an executable capability to a block of instr_length (push) + instr_length(jmp)>
+\begin{lstlisting}
+[basicstyle=\sourcecolor{}\ttfamily] 
+p1:       <assert that r1 contains an executable capability to a block of 
+           instr_length (push) + instr_length(jmp)>
           <load return capability to p1-ret1 in r2>
@@
 
-p2:       <assert that r1 contains an executable capability to a block of instr_length (push) + instr_length(jmp)>
+p2:       <assert that r1 contains an executable capability to a block of 
+           instr_length (push) + instr_length(jmp)>
           <load return capability to p2-ret1 in r2>
@@
 Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+\begin{lstlisting}
+[basicstyle=\sourcecolor{}\ttfamily] 
 closure: push r3
@@
 The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+\begin{lstlisting}
+[basicstyle=\sourcecolor{}\ttfamily] 
 evil-closure: jnz r3 (\pcreg+2)
@@
 Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
+\lau{07-09-2017: One would use $\scall{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
+The return would happen just by using $\sxjmp{}{}$ with the $\retptrd$ and $\retptrc$.
+However, this would work with the local capabilities as the real implementation of $\retptrd$ would be a sealed version of the entire stack pointer.
+In other words, upon jumping back no extra work would have to be done.
+Now, however, the old stack pointer should be spliced (using the new $\ssplice{}{}$ instruction) with the stack pointer we sent away in order to regain the old stack pointer.
+In other words, we need something to recognise when a return happens.}
 

### 2017-09-07 fully-abstract-comp/technical-report.tex
another comment

@@
 In other words, we need something to recognise when a return happens.}
+\lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{}{}$ instruction.
+The real implementation of the $\retptrc$ will simply point to a piece of clean up code that splices together the old/private stack pointer with the one we got from the untrusted party.
+This may, however, give some issues when we want to reason about it because there will be a disconnect between where the clean up job is done in the operational semantics of the source language (namely as part of the $\sxjmp{}{}$), but in a translated program, the code that does the clean up origins with the $\scall{}{}$ instruction.
+}
 

### 2017-09-08 fully-abstract-comp/technical-report.tex
Simplify an example.

@@
 The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
-\begin{lstlisting}
-[basicstyle=\sourcecolor{}\ttfamily] 
-evil-closure: jnz r3 (\pcreg+2)
+\begin{lstlisting}[basicstyle=\targetcolor{}\ttfamily] 
+evil-closure: jnz r3 r2
               store \rstk r3
@@
 \end{lstlisting}
-\dominique{7-9-2017: the above assumes that we have a way of jnz-ing to pc + 2, which may not be true in our actual assembly, but I haven't managed to construct an example that falls entirely within our language.}
 If we now assume that the instructions in evil-closure (jnz+store+jmp) are the same length as the original closure(store+increment+jmp), then the adversary can now construct a closure that discriminates $\mathtt{p1}$ and $\mathtt{p2}$ above.

### 2017-09-08 fully-abstract-comp/technical-report.tex
Add a write-up of what Lau and I discussed earlier.

@@
 This alternative approach stays even closer to the target language.
-Specifically, rather than using push/pop and call/return instructions, we do not add any instructions, but instead use special token values that play the role of the stack pointer/return pointer.
+The problem is not actually that we shouldn't offer push/pop and call/return instructions, but rather that our language should also offer an equivalent for the more low-level interactions with the stack that are possible in the target language.
+As such, we introduce special token values that play the role of the stack pointer/return pointer and we allow all interactions with them that are possible using instructions like \sstore{}{}, \smove{}{} etc.
 Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the code and data parts of the sealed return closure.
-Rather than invoking push or pop, one would store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:source-language}, this has the semantic effect that one would expect.
-Similarly, rather than invoking return, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.
-Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
-\lau{07-09-2017: One would use $\scall{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
+Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:source-language}, this has the semantic effect that one would expect.
+Similarly, rather than invoking \sreturn{}, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.
+
+We do not offer primitive \sreturn{}, \spush{} and \spop{} instructions, even though there is nothing keeping us from doing that, either as an additional primitive instruction or as a macro on top of the primitive token values mentioned above.
+We do offer a primitive \scall{}{} instruction, and it offers the only way to create a new stack frame.
+It is the only instruction that does not have a target language counterpart.
+However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{}{} compiles to.
+We will discuss later that this means that if we backtranslate an adversary piece of code (in the full abstraction proof), then it will be backtranslated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
+In other words, the back-translated code never uses call to protect its own stack frame from the code that it passes control to.
+This means that the back-translated code is quite insecure, but we don't need it to be: it just needs to be behaviorally equivalent to the target language code that it represents.
+
+% Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
+% \lau{07-09-2017: One would use $\scall{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
 The return would happen just by using $\sxjmp{}{}$ with the $\retptrd$ and $\retptrc$.
@@
 In other words, we need something to recognise when a return happens.}
-\lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{}{}$ instruction.
+% \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{}{}$ instruction.
 The real implementation of the $\retptrc$ will simply point to a piece of clean up code that splices together the old/private stack pointer with the one we got from the untrusted party.
@@
 
-Note that this weird behavior is essentially just a weird syntax for something we would otherwise write using call/return and push/pop.
-However, it removes the problem described above of relying on facts that follow from code length, because we can now write exactly the same code in source and target and it will do exactly the same thing if it is well-behaved.
+Note that the token values representing the stack and return pointer in our source language are essentially just a weird alternative syntax for something we would otherwise write using call/return and push/pop.
+However, it does remove the problem described above of relying on facts that follow from code length, because we can now write exactly the same code in source and target and it will do exactly the same thing if it is well-behaved.
 The code would look exactly the same in source and target, it would just be working with token values in the source versus real stack and return pointers in the target.
@@
 
+\subsubsection{Conditional full abstraction}
+\label{sec:conditional-full-abstraction}
+
+Another novelty in this work is the fact that we do not want to prove general full abstraction (i.e.
+two source programs are behaviorally equivalent if and only if they are compiled to behaviorally equivalent target programs).
+The reason is that we only want to prove a result about programs that respect the calling convention, i.e.\ that do not do anything stupid.
+For example, in the old setup with global/local capabilities, when trusted code performs an \scall{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
+Such sanity conditions are necessary (we cannot guarantee anything for code that does such stupid things) but we do not want to spend computation time on checking them.
+Instead, we expect a reasonable compiler to guarantee in some way that its generated code never does such stupid things and then our guarantees should apply.
+
+Note that in the new setup with linear capabilities instead of local/global ones, there seems to be less opportunity for the trusted code to do stupid things, so it may be that we no lnoger need these sanity conditions, but this is not clear yet.
+
+Assuming that we need sanity conditions on the source code, the question is how we formulate them in our full abstraction.
+Our idea is to formulate a notion of conditional full abstraction: ``if source programs t1 and t2 never perform stupid things under any possible source interaction, then they are source-equivalent iff their compilations are target-equivalent.''.
+However, the question remains how we formalise the condition ``if source programs t1 and t2 never perform stupid things under any possible source interaction''.
+
+Two possible approaches:
+\begin{itemize}
+\item \emph{Blame}: the idea here would be that we introduce a notion of ``undefined behavior'' in the operational semantics.
+  ``Doing stupid things'' like leaving copies of the stack pointer around on a call, would cause the system to enter an undefined state.
+  We would then like to express the above mentioned sanity condition as ``t1 and t2 never evaluate to undefined when plugged into an arbitrary program context''.
+  However, that doesn't work, because it does not discriminate between undefined behavior caused by t1 or t2 (on the one hand) and the program context (on the other hand).
+  To remedy this, we intended to use a notion of blame assignment: the undefined state would contain the address of the pc that was active at the moment it went wrong.
+  We would then express the above mentioned sanity condition as ``t1 and t2 never evaluate to undefined(pc) where pc is within the trusted code's assigned code memory when plugged into an arbitrary program context''.
+
+  The problem with this approach is that it does not combine well with dynamically generated code (solution: exclude code-generation?) and that it adds a lot of complexity.
+
+\item \emph{A logical relation}: An alternative approach (suggested by Lars to Lau) might be to define a logical relation to express the sanity condition ``t1 and t2 never evaluate to undefined when plugged into an arbitrary program context''.
+  This remains to be thought through in more detail, but it might be simpler and/or be easier to combine with dynamic code generation.
+\end{itemize}
+
 \end{document}

### 2017-09-08 fully-abstract-comp/technical-report.tex
added first draft for the days discussions

@@
 The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
-\begin{lstlisting}[basicstyle=\targetcolor{}\ttfamily] 
+\begin{lstlisting}
+[basicstyle=\targetcolor{}\ttfamily] 
 evil-closure: jnz r3 r2
@@
 
+\subsection{Linear Capabilities}
+The idea with linear capabilities is to limit aliasing of a capability.
+We specifically wanted a linear capability for the stack pointer, to not have to erase large pieces of memory, which we have to when we use local capabilities.
+With linear capabilities, we would just need to find the one linear capability to be sure we have the only capability under our control.
+
+There are multiple ways that linear capabilities could work (we choose the second one):
+\begin{enumerate}
+\item A linear capability comes with a system wide guarantee that no alias for this capability (or any part of the memory it has authority over).
+In other words, the linear capability is the only capability for this piece of memory.
+
+In order for this to work, the memory must initilaly be partitioned into two part.
+One controlled by a non-linear capability and one controled by a linear capability.
+The semantics of the machine would now have to make sure that the linear capability is never aliased.
+For instance, moving a linear capability from one register would cause the source register to be cleared.
+
+In this scheme, it would be okay to permit linear capabilities to become non-linear (which would consume the linear capability).
+The reasoning would be that it is okay to remove the aliasing restriction and thus make the part of memory with the no-aliasing guarantee smaller.
+If some program depends on a certain capability received from an adversary to be linear, then they would simple have to check whether it is in fact linear.
+As the linear capability is consumed when making it non-linear, it would have no way to regain the linear capability.
+
+\item A linear capability can be created from a non-linear capability.
+The linear capability has no guarantee that it is not aliased, but no aliases can be made from the linear capability.
+The linear capability can now be passed to an adversary and if the adversary gives back the linear capability when they return, then we know that they have no alias for it (provided we did not give them one and they did not have one in the first place), because the linear capability does not permit them to creat an alias.
+
+In this scheme, the system would initially have one non-linear capability for all of the memory. The linear capabilities would then be generated as they are needed.
+\end{enumerate}
+Notice that the two schemes are incompatible as they would undermine eachother.
+
+We choose to use the second scheme for the following reasons:
+\begin{enumerate}
+\item While the stack pointer was the motivation for the linear capabilities, we believe they can also be used for other things such as return pointers.
+When used for return pointers, we know that the return pointer used to return to us was the only one available.
+However, if we used the first scheme, then all of the code would have to be in the linear part of memory.
+This would be undesirable if we want to provide several entities access to the same program.
+\end{enumerate}
+
+\subsubsection{Write-linear capabilities}
+At the time is writing, it is unclear whether write-linear (in the style of write-local permission\footnote{Here it would mean that linear capabilities can only be written through capabilities with write-linear permission and all capabilities with write-linear permission must themselves be linear.}) is necessary.
+One might think that it is necessary to prevent ``indirect aliasing'', which here would mean that a linear capability would be stored in a piece of memory for which there are multiple non-linear capabilities.
+This does, however, not seem to be a problem as the aliasing invariants are never broken.
+The memory the linear capability governs is still not aliased and if anyone needs to use the capability, they would have to load it to a register which would remove it from memory.
+
+One thing write-linear capabilities may be useful for is to make sure that return capabilities are used in the correct order, i.e., ensuring well-bracketedness.
+\lau{08-09-2017: I have not thought the above completely through}
+
+We have also considered removing local capabilities as these no longer seem necessary and only add complexity.
+However, they may be relevant to ensure the well-bracketed control flow (this will need further consideration).
+
 \end{document}

### 2017-09-11 popl-2018/paper.tex
tried to fix format issues (it all looks the same, but it seems like newer template files where available)
update with low-level changes from reviews. Noted in review file what things had been addressed.

@@
 %% For double-blind review submission
-\documentclass[acmsmall,review,10pt,anonymous]{acmart}\settopmatter{printfolios=true}
+\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true}
 %% For single-blind review submission
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%\documentclass[acmsmall,10pt,review]{acmart}\settopmatter{printfolios=true}
 %% For final camera-ready submission
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+%\documentclass[acmsmall10pt,]{acmart}\settopmatter{}
 
 %% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format should change 'acmlarge' to
-%% 'sigplan,10pt'.
+%% SIGPLAN proceedings format should change 'acmsmall' to
+%% 'sigplan'.
 
@@
 
-\makeatletter
-\setlength\mpr@andskip{1em}
-\def\mpr@lineskip{5em}
-\def\MathparLineskip{\mpr@lesslineskip}
-\makeatother
+% \makeatletter
+% \setlength\mpr@andskip{1em}
+% \def\mpr@lineskip{5em}
+% \def\MathparLineskip{\mpr@lesslineskip}
+% \makeatother
 
@@
 %% Supplied to authors by publisher for camera-ready submission
-% \acmJournal{PACMPL}
-% \acmVolume{1}
-% \acmNumber{1}
-% \acmArticle{1}
-% \acmYear{2017}
-% \acmMonth{1}
-% \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-% \startPage{1}
-% \else\makeatother
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{1}
+\acmArticle{1}
+\acmYear{2017}
+\acmMonth{1}
+\acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+\else\makeatother
 %% Conference information (used by SIGPLAN proceedings format)
 %% Supplied to authors by publisher for camera-ready submission
-% \acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
-% \acmYear{2017}
-% \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-% \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-% \startPage{1}
-% \fi
+\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
+\acmYear{2017}
+\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+\acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+\fi
 
@@
 %% Title information
-\title{Reasoning about a Capability Machine with Local Capabilities}         %% [Short Title] is optional;
+\title{Reasoning About a Capability Machine with Local Capabilities}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 
-% Contrary to more traditional assembly languages, capability machines do not %
-allow literal addresses to appear as part of instructions. This means that we
+% Contrary to more traditional assembly languages, capability machines do not 
+% allow literal addresses to appear as part of instructions. This means that we
 
@@
 The logical relation helps by telling us what we need to ascertain or check in
-such scenario's to guarantee safety and preserve our invariants, but such checks
+such scenarios to guarantee safety and preserve our invariants, but such checks
 may be costly and it is not clear to us whether there are practical scenario's

### 2017-09-13 fully-abstract-comp/technical-report.tex
Removed the local capabilities (done quick - there may be things I missed).

@@
 \newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
-\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
-\newcommand{\sgetlin}[2]{\twoinstr{\src{getlin}}{#1}{#2}}
+%\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
+\newcommand{\sgetlin}[2]{\twoinstr{\src{get}}{#1}{#2}}
 \newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
@@
 \newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
-\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
-\newcommand{\tgetlin}[2]{\twoinstr{\trg{getlin}}{#1}{#2}}
+%\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
+\newcommand{\tgetlin}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
 \newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
@@
 \newcommand{\ExecConf}{\shareddom{ExecConf}}
-\newcommand{\Global}{\shareddom{Global}}
+%\newcommand{\Global}{\shareddom{Global}}
 \newcommand{\Linear}{\shareddom{Linear}}
@@
 \newcommand{\aendbnf}{\shareddom{end}}
-\newcommand{\glbnf}{\shareddom{gl}}
+%\newcommand{\glbnf}{\shareddom{gl}}
 \newcommand{\linbnf}{\shareddom{l}}
@@
 \newcommand{\perm}{\var{perm}}
-\newcommand{\gl}{\var{g}}
+%\newcommand{\gl}{\var{g}}
 \newcommand{\lin}{\var{l}}
@@
 %%% locality
-\newcommand{\plainlocality}[1]{\mathrm{#1}}
-\newcommand{\glob}{\plainlocality{global}}
-\newcommand{\local}{\plainlocality{local}}
+%\newcommand{\plainlocality}[1]{\mathrm{#1}}
+%\newcommand{\glob}{\plainlocality{global}}
+%\newcommand{\local}{\plainlocality{local}}
 
@@
 \newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
-\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
-\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
+%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
+%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
 \newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
-\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
+%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
 \newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
@@
 I do, however, not think that it is strictly necessary here to make things work.}
+\lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
 Source language domains:
@@
     \permbnf \in& \Perm & \defbnf & \dots \\
-    &\glbnf & \defbnf & \glob \mid \local \\
     &\linbnf & \defbnf & \linear \mid \normal \\
@@
     &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
-    &\scbnf &\defbnf & ((\permbnf,\glbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
+    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf}\\
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
@@
 \lau{06-09-2017: Should seals also have the option to be linear?}
+\lau{13-09-2017: For now I have added linearity to the seals.}
 The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
 \begin{itemize}
-\item $\glbnf$ defines domain $\Global$
 \item $\linbnf$ defines domain $\Linear$
@@
 \trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
-      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetloc{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
+      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
       & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
@@
 \end{figure}
-\lau{30-08-2017: We should maybe add a picture with local/global hieararchy.}
 \lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear.
@@
 It will, however, remain an invariant that there are no aliases for linear capabilities.}
-\lau{I had forgotten that return capabilities probably need to be linear.
+\lau{07-09-2017: I had forgotten that return capabilities probably need to be linear.
 This would call for a ``load from offset'' operation to be able to load an in code stored seal (see the figure I drew about linking/seals).}
+\lau{13-09-2017: Note that most of the above discussion is now captured in Section~\ref{sec:linear-cap}}
 
@@
 \[
-  \writeLocalAllowed{} \defeq \dots
-\]
-
-\[
-  % Lau: this should work for everything that might be local - used in store.
-  \isLoc{c} \defeq \dots
-\]
-
-\[
   \isLinear{\cb} \defeq \dots
@@
   \nonLinear{\cb} \defeq \dots
-\]
-
-\[
-  \nonLoc{c} \defeq \dots
 \]
@@
 \label{sec:source-language}
-\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated. It was $\sxjmp{}{}$ that needed an update.}
+\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}
 
@@
 The code capability must be non-executable and the stack register must contain a stack pointer.
-Futher the code and data capability must be local and there must be no attempt at passing on the stack capability.
 \lau{30-08-2017: Add more description here.}
 
-For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{30-08-2017: I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}
+For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee,
 \dominique{7-9-2017: We declare the call to be undefined if the caller is doing something ``stupid'' that would break security but which we do not want to check for.
@@
                                         \noRetStkMs{\ms_\stk'} \text{ and } \\
-                                        \nonLoc{\cb_1,\cb_2} \text{ and } \\
-                                        \reg' = \reg''
+                                        \reg' = \reg
                                         \begin{array}[t]{l}
@@
                                         \end{array}
-\\
-                                        \text{where $\forall r \in \RegName \ldotp \nonLoc{\reg''(r)} \text{ and } \reg''(r) = \reg(r)$}\\
-                                        \quad \text{$\text{ or } \reg''(r)=0 $} \tand \\
-                                        \text{ for $i \in \{1,2\}$ if $\isLinear{\Phi(r_i)}$, then $\reg''(r_i) = 0$ and} \\
+\\                                        \text{ for $i \in \{1,2\}$ if $\isLinear{\Phi(r_i)}$, then $\reg''(r_i) = 0$ and} \\
                                         \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
@@
                                           \noRetStkReg{\reg} \text{ and } \\
-                                          \noRetStkMs{\ms_\stk'} \text{ and } \\
-                                          \nonLoc{\cb_1,\cb_2}
+                                          \noRetStkMs{\ms_\stk'}
                                         \end{array} \right)
@@
                                         \Phi(r_2) = c \in \SealableCaps \text{ and } \\
-                                        \Phi(r_3) = \seal{\sigma,\_} \text{ and } \\
+                                        \Phi(r_3) = \seal{\lin,\sigma,\_} \text{ and } \\
                                         \vsc = \sealed{c,\sigma}
@@
                                   \Phi = (\mem, \_, \_, \_) \tand \\
-                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                  \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                   \perm \in \readAllowed{} \tand \withinBounds{\var{c}} \tand \\
@@
                                   \Phi = (\mem, \_, \_, \_) \tand \\
-                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                  \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                   \perm \in \readAllowed{} \cup \writeAllowed{} \\
@@
                                  \begin{array}[t]{l}
-                                   \Phi(r_1) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                   \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                    \perm \in \writeAllowed{} \tand \withinBounds{\var{c}} \tand \\
                                    \var{w} = \Phi(r_2) \tand \\
-                                   \isLinear{\var{w}} \tand \\
-                                   \text{if } \var{w} = ((\_,\local,\_),\_,\_,\_) \\
-                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                   \isLinear{\var{w}}
                                  \end{array} \\
@@
                                  \begin{array}[t]{l}
-                                   \Phi(r_1) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
+                                   \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                    \perm \in \writeAllowed{} \tand \withinBounds{\var{c}} \tand \\
                                    \var{w} = \Phi(r_2) \tand \\
-                                   \nonLinear{\var{w}} \tand \\
-                                   \text{if } \var{w} = ((\_,\local,\_),\_,\_,\_) \\
-                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                   \nonLinear{\var{w}}
                                  \end{array} \\
@@
                                    \var{w} = \Phi(r_2) \tand \\
-                                   \isLinear{w} \tand \\
-                                   \text{if } \isLoc{\var{w}} \text{,} \\
-                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                   \isLinear{w} 
                                  \end{array} \\
@@
                                    \var{w} = \Phi(r_2) \tand \\
-                                   \nonLinear{w} \tand \\
-                                   \text{if } \isLoc{\var{w}} \text{,} \\
-                                   \text{ then } \perm \in \writeLocalAllowed{}
+                                   \nonLinear{w}
                                  \end{array} \\
@@
 
-\textbf{Getters: geta, getb, gete, getp, getloc, and getlin}
+\textbf{Getters: geta, getb, gete, getp, getl}
 
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,\aaddr) \text{ or } \\
+                                  \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \text{ or } \\
                                   \quad \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} \tand\\
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\baddr,\_,\_) \text{ or } \\
+                                  \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \text{ or } \\
                                   \quad \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} \tand\\
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\_,\eaddr,\_) \text{ or } \\
+                                  \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \text{ or } \\
                                   \quad \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} \tand\\
@@
 
-\lau{01-09-2017: Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}
+\lau{01-09-2017: Notice, the opacity do not hide ``semantic properties'' such as permission and linearity (this could be changed if it is an issue)}
 
@@
                               \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_,\_),\_,\_,\_) \text{ or } \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}\\
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_),\_,\_,\_) \text{ or } \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}\\
                                 \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
@@
 \begin{align*}
-  \tsem{\sgetloc{r_1}{r_2}} = &                              
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\gl}) & \Phi(r_2) = ((\_,\gl,\_),\_,\_,\_) \\
-                                \updPcAddr{\Phi\updReg{r_1}{\local}} & \Phi(r_2) = \stkptr{\_,\_,\_,\_}\\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
   \tsem{\sgetlin{r_1}{r_2}} = &                              
                               \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\_,\lin),\_,\_,\_) \\
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
                                 \updPcAddr{\Phi\updReg{r_1}{\linear}} & \Phi(r_2) = \stkptr{\_,\_,\_,\_}\\
@@
 
-\textbf{Getters: geta, getb, gete, getp, getloc, and getlin}
+\textbf{Getters: geta, getb, gete, getp, and getl}
 Notice the difference between the getters in the target languages and their equivalents in the source language is that these do not have a case for the stack pointer as it does not exist in the language.
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\_,\_,\aaddr) \tand \\
+                                  \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \tand \\
                                   \perm \not\in \opaquePerm{}
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\baddr,\_,\_) \tand \\
+                                  \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \tand \\
                                   \perm \not\in \opaquePerm{}
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_,\_),\_,\eaddr,\_) \tand \\
+                                  \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \tand \\
                                   \perm \not\in \opaquePerm{}
@@
                               \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_,\_),\_,\_,\_)\\
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_),\_,\_,\_)\\
                                 \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
@@
 \begin{align*}
-  \tsem{\tgetloc{r_1}{r_2}} = &                              
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\gl}) & \Phi(r_2) = ((\_,\gl,\_),\_,\_,\_) \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
   \tsem{\tgetlin{r_1}{r_2}} = &                              
                               \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\_,\lin),\_,\_,\_) \\
+                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
                                 \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_1) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                       \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                        \perm \in \writeAllowed{} \tand \\
                                        \baddr \leq \aaddr \leq \eaddr \tand \\
-                                       \isLoc{\Phi(r_2)} \Rightarrow \perm \in \writeLocalAllowed{} \tand \\
                                        \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                  \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                   \perm \in \readAllowed{} \tand \\
@@
                                        \begin{array}[t]{l}
-                                         \Phi(r_2) = ((\perm,\gl,\lin),\baddr, \eaddr, \aaddr) \tand \\
+                                         \Phi(r_2) = ((\perm,\lin),\baddr, \eaddr, \aaddr) \tand \\
                                          \text{for $i \in \{3,4\}$ and} \\
@@
                                          \quad \perm \neq \enter \tand \\
-                                         \quad c = ((\perm,\gl,\lin),n_3,n_4,\aaddr)
+                                         \quad c = ((\perm,\lin),n_3,n_4,\aaddr)
                                        \end{array} \\
@@
                                        \begin{array}[t]{l}
-                                         \Phi(r_2) = \seal{\perm,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
+                                         \Phi(r_2) = \seal{\lin,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
                                          \text{for $i \in \{3,4\}$ and} \\
@@
                                          \quad \perm \neq \enter \tand \\
-                                         \quad c = \seal{\perm, n_3,n_4}
+                                         \quad c = \seal{\lin, n_3,n_4}
                                        \end{array} \\
@@
                                    \begin{array}[t]{l}
-                                     \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                     c = ((\perm,\gl,\lin),\baddr,\eaddr,\baddr) \\
+                                     \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                     c = ((\perm,\lin),\baddr,\eaddr,\baddr) \\
                                      \perm \neq \enter \tand \perm \not\in \opaquePerm{}
@@
                                         \Phi(r_2) \in \SealableCaps \text{ and } \\
-                                        \Phi(r_3) = \seal{\gl,\phi_{\var{base}},\phi_{\var{end}}} \\
+                                        \Phi(r_3) = \seal{\lin,\phi_{\var{base}},\phi_{\var{end}}} \\
                                         \phi_{\var{base}} \leq \phi_{\var{end}} % This allows one to restrict a seal to not have any authority
@@
                                   \begin{array}[t]{l}
-                                    \Phi(r_1) = ((\perm, \gl, \lin), \baddr_1, \eaddr_1, \aaddr_1) \\
-                                    \Phi(r_2) = ((\perm, \gl, \lin), \baddr_2, \eaddr_2, \aaddr_2) \\
+                                    \Phi(r_1) = ((\perm, \lin), \baddr_1, \eaddr_1, \aaddr_1) \\
+                                    \Phi(r_2) = ((\perm, \lin), \baddr_2, \eaddr_2, \aaddr_2) \\
                                     \eaddr_1 + 1 = \baddr_2 \\
                                     \perm \neq \enter \\
-                                    c = ((\perm, \gl, \lin), \baddr_1, eaddr_2, \aaddr_1)
+                                    c = ((\perm, \lin), \baddr_1, eaddr_2, \aaddr_1)
                                   \end{array}
@@
                                  \begin{array}[t]{l}
-                                   \Phi(r_1) = ((\perm, \gl, \lin), \baddr, eaddr, \aaddr) \\
+                                   \Phi(r_1) = ((\perm, \lin), \baddr, eaddr, \aaddr) \\
                                    r_2 = n \text{ or } \Phi(r_2) = n \text{ and in either case } n \in \nats\\
@@
                                    \baddr \leq n  < \eaddr\\
-                                   c_1 = ((\perm,\gl,\lin),\baddr, n, \aaddr) \\
-                                   c_2 = ((\perm,\gl,\lin),n + 1, \eaddr,\aaddr)
+                                   c_1 = ((\perm,\lin),\baddr, n, \aaddr) \\
+                                   c_2 = ((\perm,\lin),n + 1, \eaddr,\aaddr)
                                  \end{array} \\
@@
 \subsection{Target program linking and layout}
+\lau{13-09-2017: TODO take another look and linking and consider what the setup must be now that we don't have local capabilities.}
 For a $\scall{}{}$ to be secure after compilation, we need to use the seal mechanism.
@@
 \subsection{Linear Capabilities}
+\label{sec:linear-cap}
 The idea with linear capabilities is to limit aliasing of a capability.
@@
 \subsubsection{Write-linear capabilities}
+
 At the time is writing, it is unclear whether write-linear (in the style of write-local permission\footnote{Here it would mean that linear capabilities can only be written through capabilities with write-linear permission and all capabilities with write-linear permission must themselves be linear.}) is necessary.
@@
 \lau{08-09-2017: I have not thought the above completely through} 
+\lau{13-09-2017: Dominique wrote something about this in an e-mail.
+The splicing necessary for each call would be what is needed to make the write-linear obsolete.
+The splicing would fail in the cases where someone try to use the wrong data-capability.
+Further something like write-linear would have the same issue as write-local.
+We would rely on the fact that the write-linear permission limits the space where we might need to look for linear capabilties.
+In order to remove the desired linear capability, we would still need to search through all the places it could be which would be infeasible (like the clear memory segment instruction).}
 

### 2017-09-13 fully-abstract-comp/technical-report.tex
updated permission hieararchy to no longer include wl permission

@@
   \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (rwlx) {$\rwlx$};
-    \node[main node] (rwx) [below of=rwlx] {$\rwx$};
+    \node[main node] (rwx) {$\rwx$};
     \node[main node] (rx) [below of=rwx] {$\rx$};
@@
 
-    \node[main node] (rwlxo) [below right of=rwlx,yshift=-0.1cm] {$\rwlxo$};
-    \node[main node] (rwxo) [below of=rwlxo] {$\rwxo$};
+    \node[main node] (rwxo) [below right of=rwx,yshift=-0.1cm] {$\rwxo$};
     \node[main node] (rxo) [below of=rwxo] {$\rxo$};
@@
 
-    \node[main node] (rwl) [below left of=rwlx,xshift=-0.5cm,yshift=-0.3cm] {$\rwl$};
-    \node[main node] (rw) [below of=rwl] {$\rw$};
+    \node[main node] (rw) [below left of=rwx,xshift=-0.5cm,yshift=-0.3cm] {$\rw$};
     \node[main node] (r) [below of=rw] {$\readonly$};
@@
 
-    \node[main node] (rwlo) [below right of=rwl,yshift=-0.1cm] {$\rwlo$};
-    \node[main node] (rwo) [below of=rwlo] {$\rwo$};
+    \node[main node] (rwo) [below right of=rw,yshift=-0.1cm] {$\rwo$};
     \node[main node] (ro) [below of=rwo] {$\ro$};
@@
     \path[every node/.style={font=\sffamily\small}]
-    (rwlxo) edge (rwxo)
     (rwxo) edge (rxo)
@@
 
-    (rwl) edge (rw)
     (rw) edge (r)
@@
 
-    (rwlo) edge (rwo)
     (rwo) edge (ro)
@@
 
-    (rwlx) edge[dashed] (rwx)
     (rwx) edge[dashed] (rx)
@@
 
-    (rwlo) edge (rwlxo)
     (rwo) edge (rwxo)
@@
 
-    (rwlo) edge (rwl)
     (rwo) edge (rw)
@@
 
-    (rwl) edge (rwlx)
-    (rw) edge[dashed] (rwx)
+    (rw) edge (rwx)
     (r) edge[dashed] (rx)
@@
 
-    (rwlxo) edge (rwlx)
-    (rwxo) edge[dashed] (rwx)
+    (rwxo) edge (rwx)
     (rxo) edge[dashed] (rx)

### 2017-09-14 fully-abstract-comp/technical-report.tex
updated the text about linear capabilities to explain why we abandon the idea of turning normal capabilities into linear capabilities.

@@
 
-There are multiple ways that linear capabilities could work (we choose the second one):
+There are multiple ways that linear capabilities could work (we choose the third one):
 \begin{enumerate}
@@
 In this scheme, the system would initially have one non-linear capability for all of the memory. The linear capabilities would then be generated as they are needed.
+\item A linear capability is ``born'' with the guarantee that it is not aliased.
+This guarantee is maintained by the semantics.
+Linear capabilties cannot be converted to another kind of capability (in other words, this is option one but where linear capabilities cannot be changed).
 \end{enumerate}
-Notice that the two schemes are incompatible as they would undermine eachother.
-
-We choose to use the second scheme for the following reasons:
-\begin{enumerate}
+Notice that the two first schemes are incompatible as they would undermine eachother.
+
+We chose to use the second scheme for the following reasons:
+\begin{itemize}
 \item While the stack pointer was the motivation for the linear capabilities, we believe they can also be used for other things such as return pointers.
@@
 This would be undesirable if we want to provide several entities access to the same program.
-\end{enumerate}
+\end{itemize}
+We later abandonned the second option for the following reason:
+\begin{itemize}
+\item The initial idea with linear capabilities was to use them for stack pointers, because there would only ever be one capability for a any part of the stack.
+However, when normal capabilities can become linear, then it is only the creator of the linear capability that can be sure that no alias is created from the linear capability.
+Someone who receives a linear capability would not be able to trust the linear capability as it might have come about as a normal capability that has been aliased and later turned linear.
+
+In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from untrusted a source we do not trust.
+\end{itemize}
 

### 2017-09-14 fully-abstract-comp/technical-report.tex
Add a note.

@@
 
-In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from untrusted a source we do not trust.
+In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from an untrusted source.
+
+Note that it would be good if malloc provided a way to initially obtain a linear capability in this model.
+Otherwise, there would never be any linear capabilities other than stack pointers, which seems a waste.
 \end{itemize}

### 2017-09-14 fully-abstract-comp/technical-report.tex
Latex fix.
Some more remarks.

@@
 The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
-\begin{lstlisting}
-[basicstyle=\targetcolor{}\ttfamily] 
+\begin{lstlisting}[basicstyle=\targetcolor{}\ttfamily] 
 evil-closure: jnz r3 r2
-              store \rstk r3
+              store $\rstk$ r3
               jmp r2
@@
 \begin{enumerate}
-\item A linear capability comes with a system wide guarantee that no alias for this capability (or any part of the memory it has authority over).
-In other words, the linear capability is the only capability for this piece of memory.
-
-In order for this to work, the memory must initilaly be partitioned into two part.
+\item A linear capability comes with a system wide guarantee that no \emph{linear} alias exists for this capability (or any part of the memory it has authority over).
+In other words, the linear capability is the only \emph{linear} capability for this piece of memory.
+However, non-linear aliases may exist.
+
+In order for this to work, the memory must initially be partitioned into two parts.
 One controlled by a non-linear capability and one controled by a linear capability.
 The semantics of the machine would now have to make sure that the linear capability is never aliased.
-For instance, moving a linear capability from one register would cause the source register to be cleared.
+For instance, moving a linear capability from one register would cause the source register to be cleared (similar to destructive reads in the ownership types literature).
 
@@
 The linear capability has no guarantee that it is not aliased, but no aliases can be made from the linear capability.
-The linear capability can now be passed to an adversary and if the adversary gives back the linear capability when they return, then we know that they have no alias for it (provided we did not give them one and they did not have one in the first place), because the linear capability does not permit them to creat an alias.
+The linear capability can now be passed to an adversary and if the adversary gives back the linear capability when they return, then we know that they have no alias for it (provided we did not give them one and they did not have one in the first place), because the linear capability does not permit them to create an alias.
 
 In this scheme, the system would initially have one non-linear capability for all of the memory. The linear capabilities would then be generated as they are needed.
-\item A linear capability is ``born'' with the guarantee that it is not aliased.
+\item A linear capability is ``born'' with the guarantee that it is not aliased, either by linear or non-linear capabilities.
 This guarantee is maintained by the semantics.
-Linear capabilties cannot be converted to another kind of capability (in other words, this is option one but where linear capabilities cannot be changed).
+Linear capabilties cannot be converted to another kind of capability (in other words, this is option one but where linear capabilities cannot be made non-linear).
+
+Note that in this scheme, it would be good if malloc provided a way to initially obtain a linear capability in this model.
+Otherwise, there would never be any linear capabilities other than stack pointers, which seems a waste.
 \end{enumerate}
@@
 \end{itemize}
-We later abandonned the second option for the following reason:
+We later abandoned the second option for the following reason:
 \begin{itemize}
@@
 In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from an untrusted source.
-
-Note that it would be good if malloc provided a way to initially obtain a linear capability in this model.
-Otherwise, there would never be any linear capabilities other than stack pointers, which seems a waste.
 \end{itemize}
@@
 
+\dominique{14-9-2017: Currently, we have removed local capabilities and we believe write-linear will (a) not work (because it would require memory clearing like write-local) and (b) not be needed (because of the splicing discussed above).}
+
 \end{document}

### 2017-09-14 fully-abstract-comp/technical-report.tex
removed opaque capabilities, added a small section about them, and added part of the unified semantics

--- 2017-09-14 fully-abstract-comp/technical-report.tex
@@
 \newcommand{\tand}{\text{ and }}
+\newcommand{\tor}{\text{ or }}
 \newcommand{\totherwise}{\text{otherwise }}
@@
 \newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
-\newcommand{\ssplice}[2]{\twoinstr{\src{splice}}{#1}{#2}}
-\newcommand{\ssplit}[2]{\twoinstr{\src{split}}{#1}{#2}}
 
@@
 \newcommand{\scseal}[3]{\threeinstr{\src{cseal}}{#1}{#2}{#3}}
+\newcommand{\ssplice}[3]{\threeinstr{\src{splice}}{#1}{#2}{#3}}
+\newcommand{\ssplit}[3]{\threeinstr{\src{split}}{#1}{#2}{#3}}
 
@@
 \newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
-\newcommand{\tsplice}[2]{\twoinstr{\trg{splice}}{#1}{#2}}
-\newcommand{\tsplit}[2]{\twoinstr{\trg{split}}{#1}{#2}}
 
 % Three arguments
+\newcommand{\tsplice}[3]{\threeinstr{\trg{splice}}{#1}{#2}{#3}}
+\newcommand{\tsplit}[3]{\threeinstr{\trg{split}}{#1}{#2}{#3}}
 \newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
@@
 \newcommand{\plainperm}[1]{\mathrm{#1}}
-\newcommand{\rwlx}{\plainperm{rwlx}}
+%\newcommand{\rwlx}{\plainperm{rwlx}}
 \newcommand{\rwx}{\plainperm{rwx}}
 \newcommand{\rx}{\plainperm{rx}}
-\newcommand{\rwlxo}{\plainperm{rwlxo}}
-\newcommand{\rwlo}{\plainperm{rwlo}}
-\newcommand{\rwl}{\plainperm{rwl}}
-\newcommand{\rwxo}{\plainperm{rwxo}}
-\newcommand{\rwo}{\plainperm{rwo}}
+%\newcommand{\rwlxo}{\plainperm{rwlxo}}
+%\newcommand{\rwlo}{\plainperm{rwlo}}
+%\newcommand{\rwl}{\plainperm{rwl}}
+%\newcommand{\rwxo}{\plainperm{rwxo}}
+%\newcommand{\rwo}{\plainperm{rwo}}
 \newcommand{\rw}{\plainperm{rw}}
-\newcommand{\rxo}{\plainperm{rxo}}
-\newcommand{\ro}{\plainperm{ro}}
+%\newcommand{\rxo}{\plainperm{rxo}}
+%\newcommand{\ro}{\plainperm{ro}}
 \newcommand{\readonly}{\plainperm{r}}
 \newcommand{\noperm}{\plainperm{0}}
-\newcommand{\nopermo}{\plainperm{0o}}
+%\newcommand{\nopermo}{\plainperm{0o}}
 \newcommand{\enter}{\plainperm{e}}
-\newcommand{\entero}{\plainperm{eo}}
+%\newcommand{\entero}{\plainperm{eo}}
 
@@
 \newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
-\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
+%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
 \newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
@@
       & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
-      & & \tsplice{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}
+      & & \tsplice{\trg{r}}{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}{\trg{r}}
 \end{array}
@@
 
-    \node[main node] (rwxo) [below right of=rwx,yshift=-0.1cm] {$\rwxo$};
-    \node[main node] (rxo) [below of=rwxo] {$\rxo$};
-    \node[main node] (eo) [below of=rxo] {$\entero$};
-
     \node[main node] (rw) [below left of=rwx,xshift=-0.5cm,yshift=-0.3cm] {$\rw$};
@@
 
-    \node[main node] (rwo) [below right of=rw,yshift=-0.1cm] {$\rwo$};
-    \node[main node] (ro) [below of=rwo] {$\ro$};
-    \node[main node] (0o) [below of=ro] {$\nopermo$};
-
     \path[every node/.style={font=\sffamily\small}]
-    (rwxo) edge (rxo)
-    (rxo) edge (eo)
 
@@
 
-    (rwo) edge (ro)
-    (ro) edge (0o)
-
-    (rwx) edge[dashed] (rx)
-    (rx) edge[dashed] (e)
-
-    (rwo) edge (rwxo)
-    (ro) edge (rxo)
-    (0o) edge (eo)
-
-    (rwo) edge (rw)
-    (ro) edge (r)
-    (0o) edge (0)
+    (rwx) edge (rx)
+    (rx) edge (e)
 
     (rw) edge (rwx)
-    (r) edge[dashed] (rx)
-    (0) edge[dashed] (e)
-
-    (rwxo) edge (rwx)
-    (rxo) edge[dashed] (rx)
-    (eo) edge[dashed] (e);
+    (r) edge (rx)
+    (0) edge (e);
   \end{tikzpicture}
@@
 \begin{itemize}
-\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
+%\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
 \item Linear capabilities are cleared when they move around in memory.
@@
   \begin{itemize}
-  \item Opaque capabilities do not hide the length of a program (especially an issue if we ever hope to have error recovery).
+%  \item Opaque capabilities do not hide the length of a program (especially an issue if we ever hope to have error recovery).
   \item This is needed for correctness. See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
@@
 \[
-  \updPcPerm{c} \defeq \dots
+  \updPcPerm{c} \defeq
+  \begin{cases}
+    ((\rx,\lin),\baddr,\eaddr,\aaddr) & c = ((\enter,\lin),\baddr,\eaddr,\aaddr)\\
+    c & \totherwise
+  \end{cases}
 \]
@@
 \[
-  \isLinear{\cb} \defeq \dots
+  \isLinear{\cb} \defeq
+  \begin{cases}
+    \top & 
+    \arraycolsep=0pt
+    \begin{array}[t]{l}
+      \cb = ((\_,\linear),\_,\_,\_) \tor\\
+      \cb = \seal{\linear,\_,\_} \tor\\
+      \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
+    \end{array}\\
+    \src{\top} & 
+    \arraycolsep=0pt
+    \begin{array}[t]{l}
+      \src{\cb = \stkptr{\_,\_,\_,\_}}\\
+      \src{\cb = \retptrd{}}
+    \end{array}\\
+    \bot & \totherwise
+  \end{cases}
 \]
@@
 \[
-  \nonLinear{\cb} \defeq \dots
+  \nonLinear{\cb} \defeq \neg \isLinear{\cb}
 \]
@@
 \[
-  \opaquePerm{} \defeq \dots
-\]
-
-\[
   \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rret, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
@@
 \[
-  \nonZero{c} \defeq \dots
+  \nonZero{w} \defeq
+  \begin{cases}
+    \bot & w \in \ints \tand w = 0 \\
+    \top & \totherwise
+  \end{cases}
 \]
+
+\subsubsection{The Operational Semantics}
+We have unified the two languages in the below definitions.
+Everything written in black is common for both source and target language.
+Everything written in \src{blue} is specific to the source language.
+
+\noindent\textbf{fail and halt}\\
+\begin{align*}
+  \sem{\tfail}(\Phi) = & \; \failed \\
+  \sem{\thalt}(\Phi) = & \; (\halted, \Phi.\mem)
+\end{align*}
+
+\noindent\textbf{jmp and jnz}\\
+\begin{align*}
+  \sem{\tjmp{r}}(\Phi) = &  
+                     \begin{cases}
+                       \Phi\updReg{r}{0}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & \isLinear{\Phi(r)} \\
+                       \Phi\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & \totherwise
+                     \end{cases}
+\end{align*}
+
+\begin{align*}
+  \sem{\tjnz{r_1}{r_2}}(\Phi) = &       
+                             \begin{cases}
+                               \Phi\updReg{\pcreg}{r_1} & \text{if $\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
+                               \Phi\updReg{r_1}{0}\updReg{\pcreg}{\Phi(r_1)}
+                                               & \arraycolsep=0pt                               
+                                               \begin{array}[t]{l}
+                                                 \isLinear{\Phi(r_1)} \tand
+                                                 \nonZero{\Phi{r_2}}
+                                               \end{array}
+\\
+                               \updPcAddr{\Phi} & \totherwise
+                             \end{cases}
+\end{align*}
+
+\noindent\textbf{gettype}\\
+In the definitions of the semantics below, we use a function $\encType{} : \Word \rightarrow \ints$.
+This is an encoding function for which the specific implementation does not matter.
+As the words of the two machines differ, we really need two functions which we call $\encType{}_{\var{src}}$ and $\encType{}_{\var{trg}}$.
+These two functions need to be related in the following way:
+\lau{14-09-2017: Insert the relation.
+My impression is that the the words specific to the source machine should have the same "type" as their implementaion at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation).
+}
+\begin{align*}
+  \sem{\tisptr{r_1}{r_2}}(\Phi) = & \; \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
+\end{align*}
+
+\noindent\textbf{geta, getb, gete, getp, and getl}\\
+We assume function to encode and decode permissions as well as a function to encode and decode linearity.
+The functions are used implicitely when a permission or linearity is used in a place where they need to be a word.
+\lau{14-09-2017: Write more about this (types and assumptions)}
+\begin{align*}
+  \sem{\tgeta{r_1}{r_2}}(\Phi) = & 
+                                   \begin{cases}
+                                     \updPcAddr{\Phi\updReg{r_1}{\aaddr}} & 
+                                     \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} } \\
+                                     \end{array} \\
+                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                                   \end{cases}
+\end{align*}
+
+\begin{align*}
+  \sem{\tgetb{r_1}{r_2}}(\Phi) = & 
+                                   \begin{cases}
+                                     \updPcAddr{\Phi\updReg{r_1}{\baddr}} & 
+                                     \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} } \\
+                                     \end{array} \\
+                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                                   \end{cases}
+\end{align*}
+
+\begin{align*}
+  \sem{\tgete{r_1}{r_2}}(\Phi) = & 
+                                   \begin{cases}
+                                     \updPcAddr{\Phi\updReg{r_1}{\eaddr}} & 
+                                     \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} } \\
+                                     \end{array} \\
+                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                                   \end{cases}
+\end{align*}
+
+\begin{align*}
+  \sem{\tgetp{r_1}{r_2}}(\Phi) = & 
+                                   \begin{cases}
+                                     \updPcAddr{\Phi\updReg{r_1}{\perm}} & 
+                                     \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \Phi(r_2) = ((\perm,\_),\_,\_,\_) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\_,\_} } \\
+                                     \end{array} \\
+                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                                   \end{cases}
+\end{align*}
+
+\begin{align*}
+  \sem{\tgetlin{r_1}{r_2}}(\Phi) = & 
+                                   \begin{cases}
+                                     \updPcAddr{\Phi\updReg{r_1}{\lin}} & 
+                                     \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \\
+                                     \end{array} \\
+                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                                   \end{cases}
+\end{align*}
+
+\noindent\textbf{move}\\
+\begin{align*}
+  \sem{\tmove{r_1}{r_2}}(\Phi) = & 
+                              \begin{cases}
+                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
+                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
+                              \end{cases}
+\end{align*}
+
+\noindent\textbf{call}\\
+\begin{align*}
+  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & \dots
+\end{align*}
+
+\noindent\textbf{cseal}\\
+\begin{align*}
+  \sem{\tcseal{r_1}{r_2}{r_3}}(\Phi) = &
+                                  \begin{cases}
+                                    \Phi\updReg{r_1}{\vsc} &
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \Phi(r_2) \in \SealableCaps \tand \\
+                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr} \tand \\
+                                      \sigma_\baddr \leq \sigma_\eaddr \tand \\ 
+                                      \vsc = \sealed{c,\sigma}
+                                    \end{array}
+                                    \\
+                                    \failed & \totherwise
+                                  \end{cases}
+\end{align*}
+
+\textbf{split}\\
+\lau{14-09-2017: Do we want split to work on seals (not sealed capabilities, but seals?)}
+\begin{align*}
+  \tsem{\tsplit{r_1}{r_2}{r_3}} = &
+                               \begin{cases}
+                                 \Phi\updReg{r_1}{\dots} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \dots
+                                 \end{array}\\
+                                 \failed & \totherwise
+                               \end{cases}
+\end{align*}
 
@@
 \label{sec:source-language}
+\lau{14-09-2017: This section will be obsolete when the unifyed definitions are done.}
 \lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}
@@
                                   \quad \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} \tand\\
-                                  \perm \not\in \opaquePerm{}
                                 \end{array} \\
@@
                                   \quad \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} \tand\\
-                                  \perm \not\in \opaquePerm{}
                                 \end{array} \\
@@
                                   \quad \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} \tand\\
-                                  \perm \not\in \opaquePerm{}
                                 \end{array} \\
@@
                         \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_r)}
-                             & \nonLinear{\Phi(r)} \\
+                             & \isLinear{\Phi(r)} \\
                         \Phi\updReg{\pcreg}{\Phi(r)} & \totherwise
@@
                                   \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \tand \\
-                                  \perm \not\in \opaquePerm{}
                                 \end{array} \\
@@
                                   \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \tand \\
-                                  \perm \not\in \opaquePerm{}
                                 \end{array} \\
@@
                                   \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \tand \\
-                                  \perm \not\in \opaquePerm{}
                                 \end{array} \\
@@
                                      c = ((\perm,\lin),\baddr,\eaddr,\baddr) \\
-                                     \perm \neq \enter \tand \perm \not\in \opaquePerm{}
                                    \end{array}
@@
 
-$\tsplice{}{}$ we allow two capabilities to be spliced if the second one starts on the address after the first one ends.
+$\tsplice{}{}{}$ we allow two capabilities to be spliced if the second one starts on the address after the first one ends.
 They must also agree on permission, globality and linearity (so we also allow non-linear capabilities to be spliced).
@@
 \begin{align*}
-  \tsem{\tsplice{r_1}{r_2}} = &
+  \tsem{\tsplice{r_1}{r_2}{r_3}} = &
                                 \begin{cases}
@@
 
-$\tsplit{}{}$ this instruction splits a capability in two.
+$\tsplit{}{}{}$ this instruction splits a capability in two.
 We require the address we wish to split on to be in the interval of the authority granted by the capability to prevent handing out authority from capabilities with no authority.
@@
 \begin{align*}
-  \tsem{\tsplit{r_1}{r_2}} = &
+  \tsem{\tsplit{r_1}{r_2}{r_3}} = &
                                \begin{cases}
@@
 
+\subsection{Opaque capabilities}
+At some point, we contemplated having opaque capabilities.
+Opaque capabilities would hide the addresses of the capability, i.e, a getter on the base, end or address would yield nothing.
+It would still be possible to manipulate opaque capabilities.
+To this end, it would be necessary to have operations that work relatively to the capabilities addresses.
+These could be:
+\begin{enumerate}
+\item reset current address of capability to its base address
+\item return difference between current address and base address of a capability
+\item increase/decrease current address of a capability by a set amount
+\end{enumerate}
+(Linear capabilities may still call for some of these to be added for practical purposes. (like when the executing capability is linear, then it cannot be moved from the pc register).)
+
+At the moment of writing this, I am not sure why we introduced them, but it was related to hiding program lengths (and stack height).
+At this juncture, it does, however, not seem necessary to have opaque because sealed capabilities can do their job.
+
 \end{document}

### 2017-09-14 fully-abstract-comp/technical-report.tex
added more op sem.

@@
                                 \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
+                              \end{cases}
+\end{align*}
+Notice that in the case where we are moving a linear capability and $r_1 = r_2$ the order of the updates matter.
+
+\noindent\textbf{store}\\
+\begin{align*}
+  \sem{\tstore{r_1}{r_2}}(\Phi) = &
+                                    \begin{cases}
+                                      \updPcAddr{\Phi\update{\mem.a}{w}}\update{\reg.r_2}{w_2} & 
+                                      \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                        \perm \in \writeAllowed{} \tand \withinBounds{\Phi(r_1)} \tand \\
+                                        w = \Phi(r_2) \tand \\
+                                        \isLinear{w} \tand w_2 = 0 \tor \\
+                                        \nonLinear{w} \tand w_2 = w
+                                      \end{array}
+                                      \\
+                                      \src{\updPcAddr{\Phi\update{\mem_\stk.a}{w}}\update{\reg.r_2}{w_2}} & 
+                                      \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        \src{\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
+                                        \src{\perm \in \writeAllowed{} \tand \withinBounds{\Phi(r_1)} \tand} \\
+                                        \src{w = \Phi(r_2) \tand} \\
+                                        \src{\isLinear{w} \tand w_2 = 0 \tor} \\
+                                        \src{\nonLinear{w} \tand w_2 = w}
+                                      \end{array}\\
+                                      \failed & \totherwise
+                                    \end{cases}
+\end{align*}
+
+\begin{align*}
+  \sem{\tload{r_1}{r_2}}(\Phi) = & 
+                                  \begin{cases}
+                                    \updPcAddr{} & \dots\\
+                                    \failed & \totherwise                                    
                                   \end{cases}
@@
 
-In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from untrusted a source we do not trust.
+In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from an untrusted source.
 \end{itemize}
@@
 However, they may be relevant to ensure the well-bracketed control flow (this will need further consideration).
+
+\dominique{14-9-2017: Currently, we have removed local capabilities and we believe write-linear will (a) not work (because it would require memory clearing like write-local) and (b) not be needed (because of the splicing discussed above).}
 

### 2017-09-15 fully-abstract-comp/technical-report.tex
moved optional arguments to lstlisting to the correct position (fixed a local auctex problem, so optional arguments no longer cause an issue)

@@
 The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
-\begin{lstlisting}
-[basicstyle=\sourcecolor{}\ttfamily] 
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
 p1 ::= if r1 is length 2 then
@@
 The context with $r1$ as an executable capability pointing to:
-\begin{lstlisting}
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
 $\scall{r5}{0}$
@@
 Imagine two trusted blocks of code that looks as follows:
-\begin{lstlisting}
-[basicstyle=\sourcecolor{}\ttfamily] 
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
 p1:       <assert that r1 contains an executable capability to a block of 
@@
 Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
-\begin{lstlisting}
-[basicstyle=\sourcecolor{}\ttfamily] 
+\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
 closure: push r3

### 2017-09-15 fully-abstract-comp/technical-report.tex
Added load, took a pass over the semantics so far, changed lea to sca (set current address) so the name reflects what the instruction does.

@@
 
-\newcommand{\zinstr}[1]{#1}
+\newcommand{\zinstr}[1]{\texttt{#1}}
 \newcommand{\oneinstr}[2]{
@@
 \newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
-\newcommand{\slea}[2]{\twoinstr{\src{lea}}{#1}{#2}}
+\newcommand{\ssca}[2]{\twoinstr{\src{sca}}{#1}{#2}}
 \newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
@@
 \newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
-\newcommand{\tlea}[2]{\twoinstr{\trg{lea}}{#1}{#2}}
+\newcommand{\tsca}[2]{\twoinstr{\trg{sca}}{#1}{#2}}
 \newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tlea{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tsca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
       & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
@@
 
-\subsubsection{The Operational Semantics}
+\subsubsection{Instruction Interpretation}
 We have unified the two languages in the below definitions.
@@
                              \begin{cases}
-                               \Phi\updReg{\pcreg}{r_1} & \text{if $\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
+                               \Phi\updReg{\pcreg}{r_1} & \text{$\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
                                \Phi\updReg{r_1}{0}\updReg{\pcreg}{\Phi(r_1)}
-                                               & \arraycolsep=0pt                               
-                                               \begin{array}[t]{l}
-                                                 \isLinear{\Phi(r_1)} \tand
-                                                 \nonZero{\Phi{r_2}}
-                                               \end{array}
-\\
+                                               & \text{$\isLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$}\\
                                \updPcAddr{\Phi} & \totherwise
@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} } \\
+                                       \Phi(r_2) = ((\_,\_),\_,\_,\aaddr) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\aaddr} } \\
                                      \end{array} \\
@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} } \\
+                                       \Phi(r_2) = ((\_,\_),\baddr,\_,\_) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\baddr,\_,\_} } \\
                                      \end{array} \\
@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} } \\
+                                       \Phi(r_2) = ((\_,\_),\_,\eaddr,\_) \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\eaddr,\_} } \\
                                      \end{array} \\
@@
 \end{align*}
+\lau{15-09-2017: Should gete and getb do something for seals?}
 
@@
 \end{align*}
+\lau{15-09-2017: Should getl do something for seals?}
+\lau{15-09-2017: Do we want to allow getl and getp to work on sealed capabilities?}
 
@@
                                     \begin{cases}
-                                      \updPcAddr{\Phi\update{\mem.a}{w}}\update{\reg.r_2}{w_2} & 
+                                      \updPcAddr{\Phi\update{\mem.a}{w}\update{\reg.r_2}{w_2}} & 
                                       \arraycolsep=0pt
@@
                                         w = \Phi(r_2) \tand \\
-                                        \isLinear{w} \tand w_2 = 0 \tor \\
-                                        \nonLinear{w} \tand w_2 = w
+                                        (\isLinear{w} \rightarrow w_2 = 0) \tand \\
+                                        (\nonLinear{w} \rightarrow w_2 = w)
                                       \end{array}
                                       \\
-                                      \src{\updPcAddr{\Phi\update{\mem_\stk.a}{w}}\update{\reg.r_2}{w_2}} & 
+                                      \src{\updPcAddr{\Phi\update{\mem_\stk.a}{w}\update{\reg.r_2}{w_2}}} & 
                                       \arraycolsep=0pt
@@
                                         \src{\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                        \src{\perm \in \writeAllowed{} \tand \withinBounds{\Phi(r_1)} \tand} \\
+                                        \src{\perm \in \writeAllowed{} \tand} \\
+                                        \src{\withinBounds{\Phi(r_1)} \tand} \\
                                         \src{w = \Phi(r_2) \tand} \\
-                                        \src{\isLinear{w} \tand w_2 = 0 \tor} \\
-                                        \src{\nonLinear{w} \tand w_2 = w}
+                                        \src{(\isLinear{w} \Rightarrow w_2 = 0) \tand} \\
+                                        \src{(\nonLinear{w} \Rightarrow w_2 = w)}
                                       \end{array}\\
@@
 
+\noindent\textbf{load}\\
 \begin{align*}
@@
                                   \begin{cases}
-                                    \updPcAddr{} & \dots\\
+                                    \updPcAddr{\Phi\updReg{r_1}{w}\update{\mem.a}{w_2}} & 
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                      \perm \in \readAllowed{} \tand\\
+                                      \withinBounds{\Phi(r_2)} \tand \\
+                                      \Phi.\mem(a) = w \tand \\
+                                      (\isLinear{w} \Rightarrow w_2 = 0) \tand \\
+                                      (\nonLinear{w} \Rightarrow w_2 = \Phi(r_2))
+                                    \end{array}
+                                    \\
+                                    \src{\updPcAddr{\Phi\updReg{r_1}{w}\update{\mem_\stk.a}{w_2}}} & 
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
+                                      \src{\perm \in \readAllowed{} \tand} \\
+                                      \src{\withinBounds{\Phi(r_2)} \tand} \\
+                                      \src{\aaddr \in \dom(\Phi.\ms_\stk) \tand} \\
+                                      \src{\Phi.\ms_\stk(a) = w \tand} \\
+                                      \src{(\isLinear{w} \Rightarrow w_2 = 0) \tand} \\
+                                      \src{(\nonLinear{w} \Rightarrow w_2 = \Phi(r_2))}
+                                    \end{array}
+                                    \\
+                                    \failed & \totherwise                                    
+                                  \end{cases}
+\end{align*}
+
+\noindent\textbf{sca}\\
+\emph{Set Current Address}
+\begin{align*}
+  \sem{\tsca{r_1}{r_2}}(\Phi) = & 
+                                  \begin{cases}
+                                    \dots & \dots \\
                                     \failed & \totherwise
@@
 \begin{align*}
-  \tsem{\tlea{r_1}{r_2}} = & \dots
+  \tsem{\tsca{r_1}{r_2}} = & \dots
 \end{align*}

### 2017-09-15 fully-abstract-comp/technical-report.tex
changed the name of lea again to cca (change current address). Added implmentation of cca, restrict and subseg.

@@
 \newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
-\newcommand{\ssca}[2]{\twoinstr{\src{sca}}{#1}{#2}}
+\newcommand{\scca}[2]{\twoinstr{\src{cca}}{#1}{#2}}
 \newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
@@
 \newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
-\newcommand{\tsca}[2]{\twoinstr{\trg{sca}}{#1}{#2}}
+\newcommand{\tcca}[2]{\twoinstr{\trg{cca}}{#1}{#2}}
 \newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
@@
 %%% Permissions
-\newcommand{\plainperm}[1]{\mathrm{#1}}
+\newcommand{\plainperm}[1]{\textsc{#1}}
 %\newcommand{\rwlx}{\plainperm{rwlx}}
@@
 \newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
-\newcommand{\encPerm}[1]{\plainfun{encocePermission}{#1}}
+\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
+\newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
 \newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tsca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
       & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
@@
 
+We assume functions $\decPerm{}$ and $\encPerm{}$.
+\lau{15-09-2017: TOOD, write more about these. Should be like in the local cap setting.}
+
 \subsection{Operational Semantics}
@@
                                     \begin{cases}
-                                      \updPcAddr{\Phi\update{\mem.a}{w}\update{\reg.r_2}{w_2}} & 
+                                      \updPcAddr{\Phi\update{\mem.a}{w}\updReg{r_2}{w_2}} & 
                                       \arraycolsep=0pt
@@
                                       \\
-                                      \src{\updPcAddr{\Phi\update{\mem_\stk.a}{w}\update{\reg.r_2}{w_2}}} & 
+                                      \src{\updPcAddr{\Phi\update{\mem_\stk.a}{w}\updReg{r_2}{w_2}}} & 
                                       \arraycolsep=0pt
@@
                                         \src{\withinBounds{\Phi(r_1)} \tand} \\
+                                        \src{a \in \dom(\Phi.\ms_\stk) \tand} \\
                                         \src{w = \Phi(r_2) \tand} \\
@@
 
-\noindent\textbf{sca}\\
-\emph{Set Current Address}
-\begin{align*}
-  \sem{\tsca{r_1}{r_2}}(\Phi) = & 
-                                  \begin{cases}
-                                    \dots & \dots \\
+\noindent\textbf{cca}\\
+\emph{Change Current Address}
+\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
+\begin{align*}
+  \sem{\tcca{r_1}{r_2}}(\Phi) = & 
+                                  \begin{cases}
+                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$ and in either case $n \in \ints$ and} \\
+                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                      \perm \neq \enter \tand \\
+                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
+                                    \end{array}
+                                    \\
+                                    \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \src{\text{either $n = r_2$ or $n = \Phi(r_2)$ and in either case $n \in \ints$ and}} \\
+                                      \src{\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
+                                      \src{\perm \neq \enter \tand} \\
+                                      \src{c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}}
+                                    \end{array}
+                                    \\
+                                    \failed & \totherwise
+                                  \end{cases}
+\end{align*}
+
+\noindent\textbf{restrict}\\
+This instruction uses the $\decPerm{}$ function.
+\begin{align*}
+  \sem{\trestrict{r_1}{r_2}{r_3}} = &
+                                      \begin{cases}
+                                        \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
+                                        \arraycolsep=0pt
+                                        \begin{array}[t]{l}
+                                          \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$ and in either case $n \in \ints$ and} \\
+                                          \decPerm{n} \sqsubseteq \perm \tand \\
+                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                          \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
+                                          \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2) \tand \\
+                                        \end{array}
+                                        \\
+                                        \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
+                                        \arraycolsep=0pt
+                                        \begin{array}[t]{l}
+                                          \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
+                                          \src{\text{either $n = r_2$ or $n = \Phi.\reg(r_2)$ and in either case $n \in \ints$ and}} \\
+                                          \src{\decPerm{n} \sqsubseteq \perm \tand} \\
+                                          \src{c = \stkptr{\decPerm{n},\baddr,\eaddr,\aaddr}}
+                                        \end{array}
+                                        \\
+                                        \failed & \totherwise
+                                      \end{cases}
+\end{align*}
+
+\noindent\textbf{subseg}\\
+\begin{align*}
+  \sem{\tsubseg{r_1}{r_2}{r_3}{r_4}}(\Phi) = &
+                                               \begin{cases}
+                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \text{for $i \in \{3, 4\}$ } \\
+                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
+                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
+                                                   \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                                   \baddr \leq n_3 \tand 
+                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
+                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
+                                                   \perm \neq \enter \tand\\
+                                                   c = ((\perm,\lin),n_1,\eaddr',\aaddr)\\
+                                                   \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
+                                                   \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2) \tand
+                                                 \end{array}
+\\
+                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \text{for $i \in \{3, 4\}$ } \\
+                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
+                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
+                                                   \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
+                                                   \baddr \leq n_3 \tand 
+                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
+                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
+                                                   \perm \neq \enter \tand\\
+                                                   c = \seal{\lin,n_1,\eaddr'}\\
+                                                   \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
+                                                   \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2) \tand
+                                                 \end{array}
+\\
+                                                 \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \src{\text{for $i \in \{3, 4\}$ }} \\
+                                                   \quad \src{\text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}}\\
+                                                   \quad \src{\text{and in either case $n_i \in \ints$ and}} \\
+                                                   \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
+                                                   \src{\baddr \leq n_3 \tand }
+                                                   \src{(n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand }\\
+                                                   \src{(n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand}\\
+                                                   \src{\perm \neq \enter \tand}\\
+                                                   \src{c = \stkptr{\perm,n_1,\eaddr',\aaddr}}\\
+                                                 \end{array}\\
                                                  \failed & \totherwise 
@@
 \begin{align*}
-  \tsem{\tsca{r_1}{r_2}} = & \dots
+  \tsem{\tcca{r_1}{r_2}} = & \dots
 \end{align*}

### 2017-09-19 fully-abstract-comp/technical-report.tex
more semantics added.
more semantics added. Status: xjmp is work in progress, call is missing.
removed large parts of the outdated semantics

@@
 \newcommand{\ssplice}[3]{\threeinstr{\src{splice}}{#1}{#2}{#3}}
-\newcommand{\ssplit}[3]{\threeinstr{\src{split}}{#1}{#2}{#3}}
 
@@
 \newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
+\newcommand{\ssplit}[4]{\fourinstr{\src{split}}{#1}{#2}{#3}{#4}}
 
@@
 \newcommand{\tsplice}[3]{\threeinstr{\trg{splice}}{#1}{#2}{#3}}
-\newcommand{\tsplit}[3]{\threeinstr{\trg{split}}{#1}{#2}{#3}}
 \newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
@@
 \newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
+\newcommand{\tsplit}[4]{\fourinstr{\trg{split}}{#1}{#2}{#3}{#4}}
 
@@
 \newcommand{\data}{\var{data}}
+\newcommand{\code}{\var{code}}
 \newcommand{\priv}{\var{priv}}
@@
 \newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
+\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
 %\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
@@
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
+    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
     &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
@@
       & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
-      & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \\
-      & & \tsplice{\trg{r}}{\trg{r}}{\trg{r}} \mid  \tsplit{\trg{r}}{\trg{r}}{\trg{r}}
+  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
+      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
 \end{array}
@@
       \src{\cb = \stkptr{\_,\_,\_,\_}}\\
-      \src{\cb = \retptrd{}}
+      \src{\cb = \retptrd(\_)}
     \end{array}\\
@@
   \nonLinear{\cb} \defeq \neg \isLinear{\cb}
+\]
+
+\[
+  \linCons{w} \defeq
+  \begin{cases}
+    0 & \isLinear{w} \\
+    w & \totherwise
+  \end{cases}
 \]
@@
                      \begin{cases}
-                       \Phi\updReg{r}{0}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & \isLinear{\Phi(r)} \\
-                       \Phi\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & \totherwise
+                       \Phi\updReg{r}{w}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & w \linCons{\Phi(r)}
                      \end{cases}
@@
                              \begin{cases}
-                               \Phi\updReg{\pcreg}{r_1} & \text{$\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
-                               \Phi\updReg{r_1}{0}\updReg{\pcreg}{\Phi(r_1)}
-                                               & \text{$\isLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$}\\
+                               \Phi\updReg{r_1}{w}\updReg{\pcreg}{\Phi(r_1)}
+                                               & w = \linCons{\Phi(r_1)} \tand \nonZero{\Phi(r_2)}\\
                                \updPcAddr{\Phi} & \totherwise
@@
                                      \updPcAddr{\Phi\updReg{r_1}{\baddr}} & 
-                                     \arraycolsep=0pt
+                                     \arraycolsep=0tpf
                                      \begin{array}[t]{l}
@@
                                     \begin{cases}
-                                      \updPcAddr{\Phi\update{\mem.a}{w}\updReg{r_2}{w_2}} & 
+                                      \updPcAddr{\Phi\updReg{r_2}{w_2}\update{\mem.a}{w}} & 
                                       \arraycolsep=0pt
@@
                                         w = \Phi(r_2) \tand \\
-                                        (\isLinear{w} \rightarrow w_2 = 0) \tand \\
-                                        (\nonLinear{w} \rightarrow w_2 = w)
+                                        w_2 = \linCons{w}
                                       \end{array}
@@
                                         \src{a \in \dom(\Phi.\ms_\stk) \tand} \\
-                                        \src{w = \Phi(r_2) \tand} \\
-                                        \src{(\isLinear{w} \Rightarrow w_2 = 0) \tand} \\
-                                        \src{(\nonLinear{w} \Rightarrow w_2 = w)}
+                                        \src{w = \Phi(r_2)} \\
+                                        \src{w_2 = \linCons{w}}
                                       \end{array}\\
@@
                                       \withinBounds{\Phi(r_2)} \tand \\
-                                      \Phi.\mem(a) = w \tand \\
-                                      (\isLinear{w} \Rightarrow w_2 = 0) \tand \\
-                                      (\nonLinear{w} \Rightarrow w_2 = \Phi(r_2))
+                                      w = \Phi.\mem(a) \tand \\
+                                      w_2 = \linCons{w}
                                     \end{array}
@@
                                       \src{\aaddr \in \dom(\Phi.\ms_\stk) \tand} \\
-                                      \src{\Phi.\ms_\stk(a) = w \tand} \\
-                                      \src{(\isLinear{w} \Rightarrow w_2 = 0) \tand} \\
-                                      \src{(\nonLinear{w} \Rightarrow w_2 = \Phi(r_2))}
+                                      \src{w = \Phi.\ms_\stk(a) \tand} \\
+                                      \src{w_2 = \linCons{w}}
                                     \end{array}
@@
                                           c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                          \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
-                                          \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2) \tand \\
+                                          w_2 = \linCons{\Phi(r_2)}
                                         \end{array}
@@
                                                    \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                                   \baddr \leq n_3 \tand 
+                                                   \baddr \leq n_3 \tand \\
                                                    (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
@@
                                                    c = ((\perm,\lin),n_1,\eaddr',\aaddr)\\
-                                                   \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
-                                                   \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2) \tand
+                                                   w_2 = \linCons{\Phi(r_2)}
                                                  \end{array}
@@
                                                    c = \seal{\lin,n_1,\eaddr'}\\
-                                                   \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
-                                                   \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2) \tand
+                                                   w_2 = \linCons{\Phi(r_2)}
                                                  \end{array}
@@
                                                    \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                                   \src{\baddr \leq n_3 \tand }
+                                                   \src{\baddr \leq n_3 \tand } \\
                                                    \src{(n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand }\\
@@
 
-\noindent\textbf{call}\\
-\begin{align*}
-  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & \dots
+\noindent\textbf{lt}\\
+\begin{align*}
+  \sem{\tlt{r_0}{r_1}{r_2}}(\Phi) = &
+                                                  \begin{cases}
+                                                    \updPcAddr{\Phi\updReg{r_0}{1}} &
+                                                    \arraycolsep=0pt
+                                                    \begin{array}[t]{l}
+                                                      \text{if for $i \in \{1,2\}$}\\
+                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
+                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                      \quad\text{and $n_1 < n_2$}\\        
+                                                    \end{array}\\
+                                                    \updPcAddr{\Phi\updReg{r_0}{1}} &
+                                                    \arraycolsep=0pt
+                                                    \begin{array}[t]{l}
+                                                      \text{if for $i \in \{1,2\}$}\\
+                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
+                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                      \quad\text{and $n_1 < n_2$}\\        
+                                                    \end{array}\\
+                                                    \failed & \text{otherwise}
+                                                  \end{cases}  
+\end{align*}
+
+\noindent\textbf{plus and minus}\\
+\begin{align*}
+  \sem{\tplus{r_0}{r_1}{r_2}}(\Phi) = &
+                                                  \begin{cases}
+                                                    \updPcAddr{\Phi\updReg{r_0}{n_1+n_2}} &
+                                                    \arraycolsep=0pt
+                                                    \begin{array}[t]{l}
+                                                      \text{if for $i \in \{1,2\}$}\\
+                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
+                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                    \end{array}\\
+                                                    \failed & \text{otherwise}
+                                                  \end{cases}  
+\end{align*}
+
+\begin{align*}
+  \sem{\tminus{r_0}{r_1}{r_2}}(\Phi) = &
+                                                  \begin{cases}
+                                                    \updPcAddr{\Phi\updReg{r_0}{n_1-n_2}} &
+                                                    \arraycolsep=0pt
+                                                    \begin{array}[t]{l}
+                                                      \text{if for $i \in \{1,2\}$}\\
+                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
+                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                    \end{array}\\
+                                                    \failed & \text{otherwise}
+                                                  \end{cases}  
+\end{align*}
+
+\noindent\textbf{seta2b}\\
+\begin{align*}
+  \sem{\tsetatob{r_1}{r_2}}(\Phi) = & 
+                                \begin{cases}
+                                  \updPcAddr{}\left(
+                                    \arraycolsep=0pt
+                                    \begin{array}{rl}
+                                    \Phi&\updReg{r_2}{w_2}\\
+                                        &\updReg{r_1}{c}
+                                    \end{array} \right)
+&
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                      \perm \neq \enter \tand \\
+                                      w_2 = \linCons{\Phi(r_2)}\\
+                                      c = ((\perm,\lin),\baddr,\eaddr,\baddr)
+                                    \end{array} \\
+                                    \failed & \totherwise
+                                \end{cases}
+\end{align*}
+
+\noindent\textbf{xjmp}\\
+\lau{18-09-2017: Work in progress - remember to delete this when sem. done.}
+\begin{align*}
+  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
+                                   \begin{cases}
+                                       \begin{array}[t]{rl}
+                                         \Phi & \updReg{r_1}{w_1}\\
+                                              & \updReg{r_2}{w_2}\\
+                                              & \updReg{\pcreg}{c_\code}\\
+                                              & \updReg{\rdata}{c_\data}
+                                       \end{array}&
+                                     \begin{array}[t]{l}
+                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
+                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
+                                       w_1 = \linCons{\Phi(r_1)} \tand \\
+                                       w_2 = \linCons{\Phi(r_2)}
+                                     \end{array}
+                                     \\
+                                     \begin{array}[t]{rl}
+                                       \Phi & \updReg{r_2}{0} \\
+                                            & \updReg{\pcreg}{\dots}\\
+                                            & \updReg{\rdata}{\dots}
+                                     \end{array}&
+                                     \begin{array}[t]{l}
+                                       \Phi(r_1) = \retptrc(\sigma) \tand \\
+                                       \Phi(r_2) = \retptrd(\sigma) \tand \\
+                                       \dots
+                                     \end{array}
+                                     \\
+                                     \failed & \totherwise
+                                   \end{cases}
 \end{align*}
@@
 \textbf{split}\\
-\lau{14-09-2017: Do we want split to work on seals (not sealed capabilities, but seals?)}
-\begin{align*}
-  \tsem{\tsplit{r_1}{r_2}{r_3}} = &
-                               \begin{cases}
-                                 \Phi\updReg{r_1}{\dots} &
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \dots
+\begin{align*}
+  \tsem{\tsplit{r_1}{r_2}{r_3}{r_4}} = &
+                               \begin{cases}
+                                 \updPcAddr{\Phi\updReg{r_3}{w}\updReg{r_1}{c_1}\updReg{r_2}{c_2}} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                   \perm \neq \enter \tand \\
+                                   \Phi(r_4) = n \tor r_4 = n \text{ and in either case } n \in \nats \\
+                                   \baddr \leq n \tand n \leq \eaddr \tand \\
+                                   c_1 = ((\perm,\lin),\baddr,n,\baddr) \tand \\
+                                   c_2 = ((\perm,\lin),n+1,\eaddr,n+1) \tand \\
+                                   w = \linCons{\Phi(r_3)}
+                                 \end{array}\\
+                                 \updPcAddr{\Phi\updReg{r_3}{w}\updReg{r_1}{c_1}\updReg{r_2}{c_2}} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
+                                   \Phi(r_4) = n \tor r_4 = n \text{ and in either case } n \in \nats \\
+                                   \sigma_\baddr \leq n \tand n < \sigma_\eaddr \tand \\
+                                   c_1 = \seal{\lin,\sigma_\baddr,n} \tand \\
+                                   c_2 = \seal{\lin,n+1,\sigma_\eaddr} \tand \\
+                                   w = \linCons{\Phi(r_3)}
+                                 \end{array}\\
+                                 \src{\updPcAddr{\Phi\updReg{r_3}{0}\updReg{r_1}{c_1}\updReg{r_2}{c_2}}} &
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \src{\Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
+                                   \src{\Phi(r_4) = n \tor r_4 = n \text{ and in either case } n \in \nats} \\
+                                   \src{\baddr \leq n \tand n \leq \eaddr \tand} \\
+                                   \src{c_1 = \stkptr{\perm,\baddr,n,\baddr} \tand} \\
+                                   \src{c_2 = \stkptr{\perm,n+1,\eaddr,n+1} \tand} 
                                  \end{array}\\
@@
                                \end{cases}
+\end{align*}
+
+\noindent\textbf{splice}\\
+\begin{align*}
+  \sem{\tsplice{r_1}{r_2}{r_3}} = &
+                              \begin{cases}
+                                \updPcAddr{}\left(
+                                \arraycolsep=0pt
+                                \begin{array}{r l}
+                                  \Phi&\updReg{r_2}{w_2}\\
+                                      &\updReg{r_3}{w_3}\\
+                                      &\updReg{r_1}{c}
+                                \end{array}\right)
+&
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = ((\perm,\lin),\baddr_2,\eaddr_2,\aaddr_2) \tand \\
+                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\aaddr_3) \tand \\
+                                  \perm \neq \enter \tand \\
+                                  \baddr_2 = \eaddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
+                                  c = ((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_2) \\
+                                  w_2 = \linCons{\Phi(r_2)} \tand \\
+                                  w_3 = \linCons{\Phi(r_3)} \\
+                                \end{array}\\
+                                \updPcAddr{}\left(
+                                \arraycolsep=0pt
+                                \begin{array}{r l}
+                                  \Phi&\updReg{r_2}{w_2}\\
+                                      &\updReg{r_3}{w_3}\\
+                                      &\updReg{r_1}{c}
+                                \end{array}\right)
+&
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2}} \tand \\
+                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3}} \tand \\
+                                  \sigma_{\baddr,2} = \sigma_{\eaddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
+                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3}} \\
+                                  w_2 = \linCons{\Phi(r_2)} \tand \\
+                                  w_3 = \linCons{\Phi(r_3)} \\
+                                \end{array}\\
+                                \updPcAddr{}\left(
+                                \arraycolsep=0pt
+                                \begin{array}{r l}
+                                  \Phi&\updReg{r_2}{0}\\
+                                      &\updReg{r_3}{0}\\
+                                      &\updReg{r_1}{c}
+                                \end{array}\right)
+&
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\aaddr_2} \tand \\
+                                  \Phi(r_3) = \stkptr{\perm,\baddr_3,\eaddr_3,\aaddr_3} \tand \\
+                                  \perm \neq \enter \tand \\
+                                  \baddr_2 = \eaddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
+                                  c = \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_2} \\
+                                \end{array}\\
+                                \failed & \totherwise
+                              \end{cases}
+\end{align*}
+
+\noindent\textbf{call}\\
+\begin{align*}
+  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & \dots
 \end{align*}
@@
 
-The $\scseal{}{}{}$ instruction seals a sealable capability with a seal.
-The base of the seal is used for the seal.
-The used seal only works if it still grants authority to seal (that is the base of the seal is smaller than or equal to the end).
-\begin{align*}
-  \ssem{\scseal{r_1}{r_2}{r_3}} = & 
-                                    \begin{cases}
-                                      \Phi\updReg{r_1}{\vsc} & 
-                                      \arraycolsep=0pt
-                                      \begin{array}[t]{l}
-                                        \Phi(r_2) = c \in \SealableCaps \text{ and } \\
-                                        \Phi(r_3) = \seal{\lin,\sigma,\_} \text{ and } \\
-                                        \vsc = \sealed{c,\sigma}
-                                      \end{array}\\
-                                      & \\
-                                      \failed & \text{otherwise}
-                                    \end{cases}
-\end{align*}
-
 Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$.
@@
 CHERI also checks whether $c_1$ is executable and in range, but we do this on every step of the operational semantics, so we do not repeat that check.
-
-\textbf{jmp}\\
-\begin{align*}
-  \ssem{\sjmp{r}} = & 
-\end{align*}
-
-\lau{01-09-2017: how should the interaction between read-only memory and linear capabilities be?
-That is, if a linear capability resides in read-only memory and you attempt to load it what happens?
-Do the load succeed with the memory cell being cleared or do the load fail (because we need to clear the cell, but we do not have the necessary permissions)?
-}
-\begin{align*}
-  \ssem{\sload{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \updPcAddr{\update{\reg.r_1}{\var{w}}} &
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi = (\mem, \_, \_, \_) \tand \\
-                                  \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
-                                  \perm \in \readAllowed{} \tand \withinBounds{\var{c}} \tand \\
-                                  \var{w} = \mem(\aaddr) \tand \\
-                                  \nonLinear{\var{w}}
-                                \end{array}\\
-                                \updPcAddr{\update{\reg.r_1}{\var{w}}\update{\mem(\aaddr)}{0}} &
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi = (\mem, \_, \_, \_) \tand \\
-                                  \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
-                                  \perm \in \readAllowed{} \cup \writeAllowed{} \\
-                                  \tand \withinBounds{\var{c}} \tand \\
-                                  \var{w} = \mem(\aaddr) \tand \\
-                                  \isLinear{\var{w}}
-                                \end{array}\\
-                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}\update{\ms_\stk(\aaddr)}{0}} & 
-                                % linear capability
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \begin{array}[t]{l}
-                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
-                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                  \perm \in \readAllowed{} \cup \writeAllowed{} \tand \\
-                                  \baddr \leq \aaddr \leq \eaddr \tand \var{w} = \ms_\stk(\aaddr) \tand \\
-                                  \isLinear{\var{w}}
-                                \end{array}
-                                \\
-                                \end{array}
-                                \\
-                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}} &
-                                \arraycolsep=0pt
-                                % non-linear capability
-                                \begin{array}[t]{l}
-                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
-                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                  \perm \in \readAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand\\
-                                  \var{w} = \ms_\stk(\aaddr) \tand \\
-                                  \nonLinear{\var{w}}
-                                \end{array}\\
-                                \failed & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \ssem{\sstore{r_1}{r_2}} = & 
-                               \begin{cases}
-                                 \updPcAddr{\Phi\update{\ms.a}{w}}\update{\reg.r_2}{0} &
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
-                                   \perm \in \writeAllowed{} \tand \withinBounds{\var{c}} \tand \\
-                                   \var{w} = \Phi(r_2) \tand \\
-                                   \isLinear{\var{w}}
-                                 \end{array} \\
-                                 \updPcAddr{\Phi\update{\ms.a}{w}} &
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
-                                   \perm \in \writeAllowed{} \tand \withinBounds{\var{c}} \tand \\
-                                   \var{w} = \Phi(r_2) \tand \\
-                                   \nonLinear{\var{w}}
-                                 \end{array} \\
-%
-                                 \updPcAddr{\Phi\update{\ms_\stk.a}{w}}\update{reg.r_2}{0} & 
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \Phi = (\_, \_, \_, \ms_\stk) \tand \\
-                                   \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                   \perm \in \writeAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand \\
-                                   \aaddr \in \dom(\ms_\stk) \tand \\
-                                   \var{w} = \Phi(r_2) \tand \\
-                                   \isLinear{w} 
-                                 \end{array} \\
-                                 \updPcAddr{\Phi\update{\ms_\stk.a}{w}} & 
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \Phi = (\_, \_, \_, \ms_\stk) \tand \\
-                                   \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                   \perm \in \writeAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand \\
-                                   \aaddr \in \dom(\ms_\stk) \tand \\
-                                   \var{w} = \Phi(r_2) \tand \\
-                                   \nonLinear{w}
-                                 \end{array} \\
-                                 \failed & \totherwise
-                               \end{cases}
-\end{align*}
-
-\textbf{Getters: geta, getb, gete, getp, getl}
-
-\begin{align*}
-  \tsem{\tgeta{r_1}{r_2}} = &
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\aaddr}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \text{ or } \\
-                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\_,\aaddr} \tand\\
-                                \end{array} \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\sgetb{r_1}{r_2}} = &
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\baddr}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \text{ or } \\
-                                  \quad \Phi(r_2) = \stkptr{\perm,\baddr,\_,\_} \tand\\
-                                \end{array} \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\sgete{r_1}{r_2}} = &
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\eaddr}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \text{ or } \\
-                                  \quad \Phi(r_2) = \stkptr{\perm,\_,\eaddr,\_} \tand\\
-                                \end{array} \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\lau{01-09-2017: Notice, the opacity do not hide ``semantic properties'' such as permission and linearity (this could be changed if it is an issue)}
-
-\begin{align*}
-  \tsem{\sgetp{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_),\_,\_,\_) \text{ or } \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}\\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\sgetlin{r_1}{r_2}} = &                              
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                \updPcAddr{\Phi\updReg{r_1}{\linear}} & \Phi(r_2) = \stkptr{\_,\_,\_,\_}\\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\textbf{gettype}\\
-\lau{06-09-2017: I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
-We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
-\lau{TODO: 06-09-2017: Write out what $\encType{}$ should satisfy}
-\begin{align*}
-  \tsem{\sisptr{r_1}{r_2}} = & 
-                               \begin{cases}
-                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
-                               \end{cases}
-\end{align*}
-
-\begin{itemize}
-\item $\scall{}{}$ does not do argument spilling anymore. We leave it to the caller to push the arguments for spilling to the stack and update the stack pointer accordingly.
-\end{itemize}
-
-\subsubsection{Target Language}
-\dominique{7-9-2017: It would be nice if we could share definitions for both languages.  Can we factor them in a way that allows this?}
-\lau{07-09-2017: I agree.
-I will first write out the two languages and then I will look at where they agree (and possibly find some way of unifying the definitions eventhough they don't agree entirely).}
-
-\begin{align*}
-  \tsem{\tfail} = & \failed
-\end{align*}
-
-\begin{align*}
-  \tsem{\thalt} = & \halted
-\end{align*}
-
-\textbf{jmp and jnz}\\
-Notice that the order of the updates of the configuration actually matters here.
-As it is below, a successful jump to $\pcreg$ results in a loop, but if the order of the updates to the configuration had been reversed, then the execution would fail as the $\pcreg$ register would contain 0.
-\begin{align*}
-  \tsem{\tjmp{r}} = & 
-                      \begin{cases}
-                        \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_r)}
-                             & \isLinear{\Phi(r)} \\
-                        \Phi\updReg{\pcreg}{\Phi(r)} & \totherwise
-                      \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tjnz{r_1}{r_2}} = &       
-                             \begin{cases}
-                               \Phi\updReg{\pcreg}{r_1} & \text{if $\nonLinear{\Phi(r_1)}$ and $\nonZero{\Phi(r_2)}$} \\
-                               \Phi\updReg{r}{0}\updReg{\pcreg}{\Phi(r_1)}
-                                               & \arraycolsep=0pt                               
-                                               \begin{array}[t]{l}
-                                                 \isLinear{\Phi(r)} \tand
-                                                 \nonZero{\Phi{r_2}}
-                                               \end{array}
-\\
-                               \updPcAddr{\Phi} & \totherwise
-                             \end{cases}
-\end{align*}
-
-\textbf{Getters: geta, getb, gete, getp, and getl}
-Notice the difference between the getters in the target languages and their equivalents in the source language is that these do not have a case for the stack pointer as it does not exist in the language.
-\begin{align*}
-  \tsem{\tgeta{r_1}{r_2}} = &
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\aaddr}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\_,\_,\aaddr) \tand \\
-                                \end{array} \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tgetb{r_1}{r_2}} = &
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\baddr}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\baddr,\_,\_) \tand \\
-                                \end{array} \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tgete{r_1}{r_2}} = &
-                              \begin{cases}
-                                \updPcAddr{\Phi\update{\reg.r_1}{\eaddr}} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\_),\_,\eaddr,\_) \tand \\
-                                \end{array} \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\lau{01-09-2017: Notice, the opacity do not hide ``semantic properties'' such as permission and linearity (this could be changed if it is an issue)}
-
-\begin{align*}
-  \tsem{\tgetp{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\encPerm{\perm}}) & \Phi(r_2) = ((\perm,\_),\_,\_,\_)\\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tgetlin{r_1}{r_2}} = &                              
-                              \begin{cases}
-                                \updPcAddr{}(\Phi\update{\reg.r_1}{\lin}) & \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
-                              \end{cases}
-\end{align*}
-
-\textbf{gettype}\\
-\lau{06-09-2017: I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
-We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
-\lau{TODO: 06-09-2017: Write out what $\encType{}$ should satisfy}
-\begin{align*}
-  \tsem{\tisptr{r_1}{r_2}} = & 
-                               \begin{cases}
-                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
-                               \end{cases}
-\end{align*}
-
-\textbf{move}\\
-\lau{06-09-2017: Like previously, the order of updates in the second case matters for the case where $r_1 = r_2$.}
-\begin{align*}
-  \tsem{\tmove{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
-                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
-                              \end{cases}
-\end{align*}
-\dominique{7-9-2017: add the condition that $r_1 \neq r_2$ for sanity?}
-\lau{07-09-2017: I would find it odd if the machine had to special case on that (what would it even do? Fail?). As it is now, it follows the invariant that a linear capability is never aliased.}
-
-\begin{align*}
-  \tsem{\tstore{r_1}{r_2}} = & 
-                               \begin{cases}
-                                 \Phi\updReg{r_2}{w_2}
-                                     \update{\mem.a}{\Phi(r_2)} & \arraycolsep=0pt
-                                     \begin{array}[t]{l}
-                                       \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                       \perm \in \writeAllowed{} \tand \\
-                                       \baddr \leq \aaddr \leq \eaddr \tand \\
-                                       \isLinear{\Phi(r_2)} \Rightarrow w_2 = 0 \tand \\
-                                       \nonLinear{\Phi(r_2)} \Rightarrow w_2 = \Phi(r_2)
-                                     \end{array} \\
-                                  \failed & \totherwise
-                               \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tload{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \Phi\update{\mem.a}{w_2}\updReg{r_1}{w} &
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                  \perm \in \readAllowed{} \tand \\
-                                  \baddr \leq \aaddr \leq \eaddr \tand \\
-                                  w = \Phi.\mem(a) \tand \\
-                                  \isLinear{w} \Rightarrow w_2 = 0 \tand \\
-                                  \nonLinear{w} \Rightarrow w_2 = \Phi.\mem(a)
-                                \end{array} \\
-                                \failed & \totherwise
-                              \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\tcca{r_1}{r_2}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\trestrict{r_1}{r_2}{r_3}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tsubseg{r_1}{r_2}{r_3}{r_4}} = &
-                                     \begin{cases}
-                                       \updPcAddr{\Phi\update{r_1}{\var{c}}} &
-                                       \arraycolsep=0pt
-                                       \begin{array}[t]{l}
-                                         \Phi(r_2) = ((\perm,\lin),\baddr, \eaddr, \aaddr) \tand \\
-                                         \text{for $i \in \{3,4\}$ and} \\
-                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
-                                         \quad \text{in either case $n_i \in \ints$ and} \\
-                                         \quad \baddr \leq n_3 \tand \\
-                                         \quad n_4 \leq \eaddr \text{ or } n_4 = \infty \tand \\
-                                         \quad \perm \neq \enter \tand \\
-                                         \quad c = ((\perm,\lin),n_3,n_4,\aaddr)
-                                       \end{array} \\
-                                       \updPcAddr{\Phi\update{r_1}{\var{s}}} &
-                                       \arraycolsep=0pt
-                                       \begin{array}[t]{l}
-                                         \Phi(r_2) = \seal{\lin,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
-                                         \text{for $i \in \{3,4\}$ and} \\
-                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
-                                         \quad \text{in either case $n_i \in \ints$ and} \\
-                                         \quad \phi_{\var{base}} \leq n_3 \tand \\
-                                         \quad n_4 \leq \phi_{\var{end}} \text{ or } n_4 = \infty \tand \\
-                                         \quad \perm \neq \enter \tand \\
-                                         \quad c = \seal{\lin, n_3,n_4}
-                                       \end{array} \\
-                                       \failed & \totherwise
-                                     \end{cases}
-\end{align*}
-\dominique{31-8-2017: Why do we allow applying subseg to a sealed capability?}
-\lau{6-9-2017: We do not. We allow subseg on the seals that are used to seal capabilities.}
-
-\begin{align*}
-  \tsem{\tlt{r_1}{r_2}{r_3}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tplus{r_1}{r_2}{r_3}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tminus{r_1}{r_2}{r_3}} = & \dots
-\end{align*}
-
-\begin{align*}
-  \tsem{\tsetatob{r_1}{r_2}} = & 
-                                 \begin{cases}
-                                   \updPcAddr{\Phi\update{\reg.r_1}{c}} &
-                                   \arraycolsep=0pt
-                                   \begin{array}[t]{l}
-                                     \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                     c = ((\perm,\lin),\baddr,\eaddr,\baddr) \\
-                                   \end{array}
-                                   \\
-                                   \failed & \totherwise
-                                 \end{cases}
-\end{align*}
-
-\begin{align*}
-  \tsem{\txjmp{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \Phi
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \update{\reg.\pcreg}{c_1} \\
-                                  \update{\reg.\rdata}{c_2}
-                                \end{array} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
-                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
-                                  \nonExec{c_2}
-                                \end{array}
-\\
-\failed & \text{otherwise}
-                              \end{cases}
-\end{align*}
-
-Sealing a capability: $\tcseal{}{}{}$ seals a sealable capability
-\begin{align*}
-  \tsem{\tcseal{r_1}{r_2}{r_3}} = & 
-                                    \begin{cases}
-                                      \updPcAddr{\Phi\update{\reg.r_1}{\sealed{\Phi(r_2),\phi_{\var{base}}}}} &
-                                      \arraycolsep=0pt
-                                      \begin{array}[t]{l}
-                                        \Phi(r_2) \in \SealableCaps \text{ and } \\
-                                        \Phi(r_3) = \seal{\lin,\phi_{\var{base}},\phi_{\var{end}}} \\
-                                        \phi_{\var{base}} \leq \phi_{\var{end}} % This allows one to restrict a seal to not have any authority
-                                      \end{array} \\
-                                      \failed & \text{otherwise}
-                                    \end{cases}
-\end{align*}
-
-$\tsplice{}{}{}$ we allow two capabilities to be spliced if the second one starts on the address after the first one ends.
-They must also agree on permission, globality and linearity (so we also allow non-linear capabilities to be spliced).
-We do, however, exclude enter capabilities as this may cause subtle security issues.
-\begin{align*}
-  \tsem{\tsplice{r_1}{r_2}{r_3}} = &
-                                \begin{cases}
-                                  \updPcAddr{\Phi\update{r_1}{\dots}\update{r_2}{c}} &
-                                  \arraycolsep=0pt
-                                  \begin{array}[t]{l}
-                                    \Phi(r_1) = ((\perm, \lin), \baddr_1, \eaddr_1, \aaddr_1) \\
-                                    \Phi(r_2) = ((\perm, \lin), \baddr_2, \eaddr_2, \aaddr_2) \\
-                                    \eaddr_1 + 1 = \baddr_2 \\
-                                    \perm \neq \enter \\
-                                    c = ((\perm, \lin), \baddr_1, eaddr_2, \aaddr_1)
-                                  \end{array}
-                                  \\
-                                  \failed & \totherwise
-                                \end{cases}
-\end{align*}
-
-$\tsplit{}{}{}$ this instruction splits a capability in two.
-We require the address we wish to split on to be in the interval of the authority granted by the capability to prevent handing out authority from capabilities with no authority.
-We prohibit enter capabilities to prevent subtle errors.
-\begin{align*}
-  \tsem{\tsplit{r_1}{r_2}{r_3}} = &
-                               \begin{cases}
-                                 \updPcAddr{\Phi\update{r_1}{c_1}\update{r_2}{c_2}} &
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \Phi(r_1) = ((\perm, \lin), \baddr, eaddr, \aaddr) \\
-                                   r_2 = n \text{ or } \Phi(r_2) = n \text{ and in either case } n \in \nats\\
-                                   \perm \neq \enter\\
-                                   \baddr \leq n  < \eaddr\\
-                                   c_1 = ((\perm,\lin),\baddr, n, \aaddr) \\
-                                   c_2 = ((\perm,\lin),n + 1, \eaddr,\aaddr)
-                                 \end{array} \\
-                                 \failed & \totherwise
-                               \end{cases}
-\end{align*}
 

### 2017-09-19 fully-abstract-comp/technical-report.tex
compilation issues
Added some comments about the current version of splice

@@
                                      \updPcAddr{\Phi\updReg{r_1}{\baddr}} & 
-                                     \arraycolsep=0tpf
+                                     \arraycolsep=0pt
                                      \begin{array}[t]{l}
@@
 \noindent\textbf{splice}\\
+\lau{19-09-2017: Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
+This means that if a caller tries to use a return pointer with a stack that is not immediately adjacent to the private stack, then it fails.
+(2) Splice prohibit splicing with an empty capability!
+This means that a callee cannot return an empty stack (this also means that it is impossible to make a call when all of the stack is used - this may indeed be undesirable, but without this restriction we need to handle other things).}
+\lau{19-09-2017: Because $\tsplice{}{}{}$ does not allow empty stacks, it is not ``left inverse'' to $\tsplit{}{}{}{}$ (because of the empty case).
+Intuitively, it is weird that a $\tsplit{}{}{}{}$ followed by a $\tsplice{}{}{}$ does not yeild the same capability.}
 \begin{align*}
@@
                                   \perm \neq \enter \tand \\
-                                  \baddr_2 = \eaddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
+                                  \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
                                   c = ((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_2) \\

### 2017-09-19 fully-abstract-comp/technical-report.tex
corrected a couple of splice errors
made parts of the operational semantics readable

@@
 \newcommand{\sourcecolortext}{blue}
-\newcommand{\sourcecolor}[1]{\color{blue}}
-\newcommand{\src}[1]{{\sourcecolor{} #1}}
+\newcommand{\sourcecolor}{\color{blue}}
+\newcommand{\src}[1]{{\sourcecolor #1}}
 \newcommand{\targetcolortext}{black}
@@
 \newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
-\newcommand{\sxjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
+\newcommand{\sxjmp}[2]{\twoinstr{\src{xjmp}}{#1}{#2}}
 \newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
@@
     \src{\top} & 
-    \arraycolsep=0pt
-    \begin{array}[t]{l}
-      \src{\cb = \stkptr{\_,\_,\_,\_}}\\
-      \src{\cb = \retptrd(\_)}
-    \end{array}\\
+    \sourcecolor\left.
+    \arraycolsep=0pt
+    \begin{array}[t]{l}
+      \cb = \stkptr{\_,\_,\_,\_}\\
+      \cb = \retptrd(\_)
+    \end{array}\right.\\
     \bot & \totherwise
@@
                              \begin{cases}
-                               \Phi\updReg{r_1}{w}\updReg{\pcreg}{\Phi(r_1)}
+                               \arraycolsep=0pt
+                               \begin{array}[t]{rl}
+                                 \Phi&\updReg{r_1}{w}\\
+                                     &\updReg{\pcreg}{\Phi(r_1)}
+                               \end{array}
                                                & w = \linCons{\Phi(r_1)} \tand \nonZero{\Phi(r_2)}\\
@@
                                     \begin{cases}
-                                      \updPcAddr{\Phi\updReg{r_2}{w_2}\update{\mem.a}{w}} & 
+                                      \updPcAddr{}\left(
+                                        \arraycolsep=0pt
+                                        \begin{array}{rl}
+                                          \Phi&\updReg{r_2}{w_2}\\
+                                              &\update{\mem.a}{w}
+                                        \end{array}
+\right) & 
                                       \arraycolsep=0pt
@@
                                         \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                        \perm \in \writeAllowed{} \tand \withinBounds{\Phi(r_1)} \tand \\
+                                        \perm \in \writeAllowed{} \tand\\
+                                        \withinBounds{\Phi(r_1)} \tand \\
                                         w = \Phi(r_2) \tand \\
@@
                                       \\
-                                      \src{\updPcAddr{\Phi\update{\mem_\stk.a}{w}\updReg{r_2}{w_2}}} & 
-                                      \arraycolsep=0pt
-                                      \begin{array}[t]{l}
-                                        \src{\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                        \src{\perm \in \writeAllowed{} \tand} \\
-                                        \src{\withinBounds{\Phi(r_1)} \tand} \\
-                                        \src{a \in \dom(\Phi.\ms_\stk) \tand} \\
-                                        \src{w = \Phi(r_2)} \\
-                                        \src{w_2 = \linCons{w}}
-                                      \end{array}\\
+                                      \sourcecolor\left.
+                                      \updPcAddr{}\left(
+                                      \arraycolsep=0pt
+                                      \begin{array}{rl}
+                                        \Phi&\update{\mem_\stk.a}{w}\\
+                                            &\updReg{r_2}{w_2}
+                                      \end{array}
+                                      \right) \right.& 
+                                      \sourcecolor\left.
+                                      \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                        \perm \in \writeAllowed{} \tand \\
+                                        \withinBounds{\Phi(r_1)} \tand \\
+                                        a \in \dom(\Phi.\ms_\stk) \tand \\
+                                        w = \Phi(r_2) \\
+                                        w_2 = \linCons{w}
+                                      \end{array}\right.\\
                                       \failed & \totherwise
@@
                                   \begin{cases}
-                                    \updPcAddr{\Phi\updReg{r_1}{w}\update{\mem.a}{w_2}} & 
+                                    \updPcAddr{}\left(
+                                      \arraycolsep=0pt
+                                      \begin{array}{rl}
+                                        \Phi&\updReg{r_1}{w}\\
+                                            &\update{\mem.a}{w_2}
+                                      \end{array}\right)& 
                                     \arraycolsep=0pt
@@
                                     \\
-                                    \src{\updPcAddr{\Phi\updReg{r_1}{w}\update{\mem_\stk.a}{w_2}}} & 
-                                    \arraycolsep=0pt
-                                    \begin{array}[t]{l}
-                                      \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                      \src{\perm \in \readAllowed{} \tand} \\
-                                      \src{\withinBounds{\Phi(r_2)} \tand} \\
-                                      \src{\aaddr \in \dom(\Phi.\ms_\stk) \tand} \\
-                                      \src{w = \Phi.\ms_\stk(a) \tand} \\
-                                      \src{w_2 = \linCons{w}}
-                                    \end{array}
+                                    \sourcecolor\left.
+                                    \updPcAddr{}\left(
+                                      \arraycolsep=0pt
+                                      \begin{array}{rl}
+                                        \Phi&\updReg{r_1}{w}\\
+                                            & \update{\mem_\stk.a}{w_2}
+                                      \end{array}\right)\right.
+                                    & 
+                                    \sourcecolor\left.
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                      \perm \in \readAllowed{} \tand \\
+                                      \withinBounds{\Phi(r_2)} \tand \\
+                                      \in \dom(\Phi.\ms_\stk) \tand \\
+                                      w = \Phi.\ms_\stk(a) \tand \\
+                                      w_2 = \linCons{w}
+                                    \end{array}\right.
                                     \\
@@
                                     \begin{array}[t]{l}
-                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$ and in either case $n \in \ints$ and} \\
+                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                      \quad\text{ and in either case $n \in \ints$ and} \\
                                       \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
@@
                                     \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
-                                    \arraycolsep=0pt
-                                    \begin{array}[t]{l}
-                                      \src{\text{either $n = r_2$ or $n = \Phi(r_2)$ and in either case $n \in \ints$ and}} \\
-                                      \src{\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                      \src{\perm \neq \enter \tand} \\
-                                      \src{c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}}
-                                    \end{array}
+                                    \sourcecolor\left.
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \text{either $n = r_2$ or $n = \Phi(r_2)$}\\
+                                      \quad\text{ and in either case $n \in \ints$ and} \\
+                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                      \perm \neq \enter \tand \\
+                                      c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
+                                    \end{array}\right.
                                     \\
@@
                                       \begin{cases}
-                                        \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
+                                        \updPcAddr{}\left(
+                                          \arraycolsep=0pt
+                                          \begin{array}{rl}
+                                          \Phi&\updReg{r_2}{w_2}\\
+                                              &\updReg{r_1}{c}
+                                          \end{array} \right)
+&
                                         \arraycolsep=0pt
@@
                                           \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$ and in either case $n \in \ints$ and} \\
+                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                          \quad\text{and in either case $n \in \ints$ and} \\
                                           \decPerm{n} \sqsubseteq \perm \tand \\
@@
                                         \\
-                                        \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
-                                        \arraycolsep=0pt
-                                        \begin{array}[t]{l}
-                                          \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                          \src{\text{either $n = r_2$ or $n = \Phi.\reg(r_2)$ and in either case $n \in \ints$ and}} \\
-                                          \src{\decPerm{n} \sqsubseteq \perm \tand} \\
-                                          \src{c = \stkptr{\decPerm{n},\baddr,\eaddr,\aaddr}}
-                                        \end{array}
+                                        \sourcecolor\left.
+                                        \updPcAddr{}\left(
+                                          \arraycolsep=0pt
+                                          \begin{array}{rl}
+                                          \Phi&\updReg{r_2}{0}\\
+                                                    &\updReg{r_1}{c}
+                                          \end{array}\right)\right.
+                                        &
+                                        \sourcecolor\left.
+                                        \arraycolsep=0pt
+                                        \begin{array}[t]{l}
+                                          \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                          \quad\text{and in either case $n \in \ints$ and} \\
+                                          \decPerm{n} \sqsubseteq \perm \tand \\
+                                          c = \stkptr{\decPerm{n},\baddr,\eaddr,\aaddr}
+                                        \end{array}\right.
                                         \\
@@
 \noindent\textbf{subseg}\\
+\lau{19-09-2017: TODO: I am not sure the way the end points are handled makes entirely sense. Have a look at it (maybe make a function that computes the end point?)}
 \begin{align*}
@@
                                                    \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
-                                                   \baddr \leq n_3 \tand 
+                                                   \baddr \leq n_3 \tand \\
                                                    (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
@@
                                                  \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
-                                                 \arraycolsep=0pt
-                                                 \begin{array}[t]{l}
-                                                   \src{\text{for $i \in \{3, 4\}$ }} \\
-                                                   \quad \src{\text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}}\\
-                                                   \quad \src{\text{and in either case $n_i \in \ints$ and}} \\
-                                                   \src{\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                                   \src{\baddr \leq n_3 \tand } \\
-                                                   \src{(n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand }\\
-                                                   \src{(n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand}\\
-                                                   \src{\perm \neq \enter \tand}\\
-                                                   \src{c = \stkptr{\perm,n_1,\eaddr',\aaddr}}\\
-                                                 \end{array}\\
+                                                 \sourcecolor\left.
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \text{for $i \in \{3, 4\}$ } \\
+                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
+                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
+                                                   \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                                   \baddr \leq n_3 \tand \\
+                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand\\
+                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
+                                                   \perm \neq \enter \tand\\
+                                                   c = \stkptr{\perm,n_1,\eaddr',\aaddr}\\
+                                                 \end{array}\right.\\
                                                  \failed & \totherwise 
@@
                                \begin{cases}
-                                 \updPcAddr{\Phi\updReg{r_3}{w}\updReg{r_1}{c_1}\updReg{r_2}{c_2}} &
+                                 \updPcAddr{}\left(
+                                   \arraycolsep=0pt
+                                   \begin{array}{rl}
+                                     \Phi&\updReg{r_3}{w}\\
+                                         &\updReg{r_1}{c_1}\\
+                                         &\updReg{r_2}{c_2}
+                                   \end{array}\right)
+&
                                  \arraycolsep=0pt
@@
                                    \perm \neq \enter \tand \\
-                                   \Phi(r_4) = n \tor r_4 = n \text{ and in either case } n \in \nats \\
+                                   \Phi(r_4) = n \tor r_4 = n \\
+                                   \quad\text{ and in either case } n \in \nats \\
                                    \baddr \leq n \tand n \leq \eaddr \tand \\
@@
                                  \end{array}\\
-                                 \updPcAddr{\Phi\updReg{r_3}{w}\updReg{r_1}{c_1}\updReg{r_2}{c_2}} &
+                                 \updPcAddr{} \left(
+                                 \arraycolsep=0pt
+                                 \begin{array}{rl}
+                                   \Phi&\updReg{r_3}{w}\\
+                                       &\updReg{r_1}{c_1}\\
+                                       &\updReg{r_2}{c_2}
+                                 \end{array} \right)
+&
                                  \arraycolsep=0pt
@@
                                    \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
-                                   \Phi(r_4) = n \tor r_4 = n \text{ and in either case } n \in \nats \\
+                                   \Phi(r_4) = n \tor r_4 = n \\
+                                   \quad\text{ and in either case } n \in \nats \\
                                    \sigma_\baddr \leq n \tand n < \sigma_\eaddr \tand \\
@@
                                  \end{array}\\
-                                 \src{\updPcAddr{\Phi\updReg{r_3}{0}\updReg{r_1}{c_1}\updReg{r_2}{c_2}}} &
-                                 \arraycolsep=0pt
-                                 \begin{array}[t]{l}
-                                   \src{\Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand} \\
-                                   \src{\Phi(r_4) = n \tor r_4 = n \text{ and in either case } n \in \nats} \\
-                                   \src{\baddr \leq n \tand n \leq \eaddr \tand} \\
-                                   \src{c_1 = \stkptr{\perm,\baddr,n,\baddr} \tand} \\
-                                   \src{c_2 = \stkptr{\perm,n+1,\eaddr,n+1} \tand} 
-                                 \end{array}\\
+                                   \sourcecolor\left.
+                                   \updPcAddr{}\left(
+                                   \arraycolsep=0pt
+                                   \begin{array}{rl}
+                                     \Phi&\updReg{r_3}{0}\\
+                                               &\updReg{r_1}{c_1}\\
+                                               &\updReg{r_2}{c_2}
+                                   \end{array} \right)\right.
+&
+                                 \sourcecolor\left.
+                                 \arraycolsep=0pt
+                                 \begin{array}[t]{l}
+                                   \Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                   \Phi(r_4) = n \tor r_4 = n \\
+                                   \quad\text{ and in either case } n \in \nats \\
+                                   \baddr \leq n \tand n \leq \eaddr \tand \\
+                                   c_1 = \stkptr{\perm,\baddr,n,\baddr} \tand \\
+                                   c_2 = \stkptr{\perm,n+1,\eaddr,n+1} 
+                                 \end{array} \right.\\
                                  \failed & \totherwise
@@
                                   \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
-                                  c = ((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_2) \\
+                                  c = ((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_2) \tand\\
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
@@
                                   \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3}} \tand \\
-                                  \sigma_{\baddr,2} = \sigma_{\eaddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
-                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3}} \\
+                                  \sigma_{\eaddr,2}+1 = \sigma_{\baddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand\\
+                                  \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
+                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3}} \tand \\
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
@@
                                 \end{array}\\
+                                \sourcecolor\left.
                                 \updPcAddr{}\left(
@@
                                       &\updReg{r_1}{c}
-                                \end{array}\right)
+                                \end{array}\right)\right.
 &
@@
                                   \perm \neq \enter \tand \\
-                                  \baddr_2 = \eaddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
+                                  \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
                                   c = \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_2} \\

### 2017-09-20 fully-abstract-comp/technical-report.tex
first draft of 'finished' operational semantics - some problems still need to be resolved.
fixed small errors in the opertational semantics.

@@
 \newcommand{\stk}{\var{stk}}
+\newcommand{\stkf}{\stk_{\var{frame}}}
 \newcommand{\ret}{\var{ret}}
@@
 \label{sec:syntax}
-The target machine is a simple capability machine with memory capabilities, local capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI).
+The target machine is a simple capability machine with memory capabilities and sealed capabilities\footnote{In previous work, we used enter capabilities but for this the complexity introduced by mixing writable and executable memory is difficult to handle.} (inspired by CHERI).
 The syntax of the instructions of the target machine is defined as follows:
@@
 
-The source machine is also a capability machine with memory capabilities, local capabilities and sealed capabilities.
+The source machine is also a capability machine with memory capabilities and sealed capabilities.
 Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities.
@@
 \[
-  \noRetStkReg{\reg} \defeq \forall r \in \reg \setminus \{ \pcreg, \rstk, \rret, \rdata \} \ldotp \reg(r) \neq \stkptr{\_}
-\]
-
-\[
-  \noRetStkMs{\ms} \defeq \forall \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) \neq \stkptr{\_}
-\]
-
-\[
   \withinBounds{c} \defeq \dots
@@
                      \begin{cases}
-                       \Phi\updReg{r}{w}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & w \linCons{\Phi(r)}
+                       \Phi\updReg{r}{w}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & w = \linCons{\Phi(r)}
                      \end{cases}
@@
                                        \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \tand l = \linear\\
                                      \end{array} \\
@@
                                                     \end{array}\\
-                                                    \updPcAddr{\Phi\updReg{r_0}{1}} &
+                                                    \updPcAddr{\Phi\updReg{r_0}{0}} &
                                                     \arraycolsep=0pt
@@
                                                       \quad\text{and in either case $n_i \in \ints$}\\
-                                                      \quad\text{and $n_1 < n_2$}\\        
+                                                      \quad\text{and $n_1 \not< n_2$}\\        
                                                     \end{array}\\
@@
                                    \begin{cases}
+                                     \arraycolsep=0pt
                                      \begin{array}[t]{rl}
@@
                                             & \updReg{\rdata}{c_\data}
-                                       \end{array}&
+                                     \end{array} &
                                      \begin{array}[t]{l}
@@
                                        \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
+                                       \nonExec{c_\data} \tand \\
                                        w_1 = \linCons{\Phi(r_1)} \tand \\
@@
                                      \end{array}
-                                     \\
+                                     \\&\\
+                                     \sourcecolor\left.
+                                       \arraycolsep=0pt
                                        \begin{array}[t]{rl}
-                                       \Phi & \updReg{r_2}{0} \\
-                                            & \updReg{\pcreg}{\dots}\\
-                                            & \updReg{\rdata}{\dots}
-                                     \end{array}&
+                                         \Phi' & \updReg{r_2}{0} \\
+                                               & \updReg{\pcreg}{\opc}\\
+                                               & \updReg{\rdata}{0}\\
+                                               & \updReg{\rstk}{c_\stk}
+                                       \end{array}\right.& 
+                                     \sourcecolor\left.
                                        \begin{array}[t]{l}
@@
                                          \Phi(r_2) = \retptrd(\sigma) \tand \\
-                                       \dots
-                                     \end{array}
+                                         \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_} \tand \\
+                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
+                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
+                                         c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
+                                         \dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk]\tand \\
+                                         \dom(\ms_{\stk,\priv} = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
+                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
+                                       \end{array}\right.
                                      \\
@@
 &
+                                \sourcecolor
                                 \arraycolsep=0pt
@@
 \noindent\textbf{call}\\
-\begin{align*}
-  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & \dots
+\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
+\lau{19-09-2017: In which cases do we have undefined here?}
+\begin{align*}
+  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & 
+                                               \begin{cases}
+                                                 \Phi'
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{rl}
+                                                   &\updReg{r_1}{w_1} \\
+                                                        &\updReg{r_2}{w_2} \\
+                                                        &\updReg{\pcreg}{c_1}\\
+                                                        &\updReg{\rdata}{c_2}\\
+                                                        &\updReg{\rstk}{c_\stk}\\
+                                                        &\updReg{r_{\var{ret,c}}}{\retptrc(\sigma_?)}\\
+                                                        &\updReg{r_{\var{ret,d}}}{\retptrd(\sigma_?)}
+                                                 \end{array}
+                                                 & 
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \Phi(r_1) = \sealed(c_1,\sigma) \tand \\
+                                                   \Phi(r_2) = \sealed(c_2,\sigma) \tand \\
+                                                   \nonExec{c_2} \tand\\
+                                                   \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
+                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
+                                                   \baddr_\stk \leq \aaddr_\stk \leq \eaddr_\stk \tand \\
+                                                   \dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk] \tand\\
+                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\baddr_\stk,\aaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
+                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand\\
+                                                   \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                                   \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
+                                                   c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
+                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk)\tand\\
+                                                   \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
+                                                   w_1 = \linCons{\Phi(r_1)} \tand \\
+                                                   w_2 = \linCons{\Phi(r_2)}
+                                                 \end{array}
+                                               \end{cases}
 \end{align*}

### 2017-09-20 fully-abstract-comp/technical-report.tex
updated semantics of move (forgot the case where an int is moved to a register) + typos
typo
Removed subseg instruction. Not needed with the split instruction.
Removed the remainder of the old semantics (i.e. call and xjmp).

@@
 % Four arguments
-\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
+%\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
 \newcommand{\ssplit}[4]{\fourinstr{\src{split}}{#1}{#2}{#3}{#4}}
@@
 % Four arguments
-\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
+%\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
 \newcommand{\tsplit}[4]{\fourinstr{\trg{split}}{#1}{#2}{#3}{#4}}
@@
     &\MemSeg & \defeq & \Addr \parfun \Word \\
-    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
+    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
     \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
   & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
@@
                                        \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \tand l = \linear\\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} \tand l = \linear}\\
                                      \end{array} \\
@@
                               \begin{cases}
+                                \updPcAddr{\Phi\updReg{r_1}{r_2}} & r_2 \in \ints \\
                                 \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
@@
                                         \\
-                                        \failed & \totherwise
-                                      \end{cases}
-\end{align*}
-
-\noindent\textbf{subseg}\\
-\lau{19-09-2017: TODO: I am not sure the way the end points are handled makes entirely sense. Have a look at it (maybe make a function that computes the end point?)}
-\begin{align*}
-  \sem{\tsubseg{r_1}{r_2}{r_3}{r_4}}(\Phi) = &
-                                               \begin{cases}
-                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
-                                                 \arraycolsep=0pt
-                                                 \begin{array}[t]{l}
-                                                   \text{for $i \in \{3, 4\}$ } \\
-                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
-                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
-                                                   \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                                   \baddr \leq n_3 \tand \\
-                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
-                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
-                                                   \perm \neq \enter \tand\\
-                                                   c = ((\perm,\lin),n_1,\eaddr',\aaddr)\\
-                                                   w_2 = \linCons{\Phi(r_2)}
-                                                 \end{array}
-\\
-                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
-                                                 \arraycolsep=0pt
-                                                 \begin{array}[t]{l}
-                                                   \text{for $i \in \{3, 4\}$ } \\
-                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
-                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
-                                                   \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
-                                                   \baddr \leq n_3 \tand \\
-                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
-                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
-                                                   \perm \neq \enter \tand\\
-                                                   c = \seal{\lin,n_1,\eaddr'}\\
-                                                   w_2 = \linCons{\Phi(r_2)}
-                                                 \end{array}
-\\
-                                                 \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
-                                                 \sourcecolor\left.
-                                                 \arraycolsep=0pt
-                                                 \begin{array}[t]{l}
-                                                   \text{for $i \in \{3, 4\}$ } \\
-                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
-                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
-                                                   \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                                   \baddr \leq n_3 \tand \\
-                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand\\
-                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
-                                                   \perm \neq \enter \tand\\
-                                                   c = \stkptr{\perm,n_1,\eaddr',\aaddr}\\
-                                                 \end{array}\right.\\
                                         \failed & \totherwise
@@
                                   \perm \neq \enter \tand \\
-                                  \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
+                                  \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
                                   c = \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_2} \\
@@
 \end{align*}
-
-\subsubsection{Source Language}
-\label{sec:source-language}
-\lau{14-09-2017: This section will be obsolete when the unifyed definitions are done.}
-\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}
-
-\textbf{Call}\\
-Call has three cases, a successful call, undefined and failed.
-For a call to be successful, it needs two sealed capabilities - a data and a code capability.
-The code capability must be non-executable and the stack register must contain a stack pointer.
-\lau{30-08-2017: Add more description here.}
-
-For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee,
-\dominique{7-9-2017: We declare the call to be undefined if the caller is doing something ``stupid'' that would break security but which we do not want to check for.
-This includes: leaving a copy of the stack pointer or the old return pointer in a place accessible to the callee, invoking local callbacks (see discussion in POPL2018 submission of why this is bad).
-The condition that $\cb_2 \neq \stkptr{\_,\_,\_,\_}$ is weird because this would not be the caller's fault, but the callee's.
-Further thought needed regarding correct blame assignment for this case.}
-
-If we are not in one of the two cases above, then the execution fails.
-\begin{align*}
-  \ssem{\scall{r_1}{r_2}} = & 
-    \begin{cases}
-      (\mem,\reg',\stk',\ms_\stk') & \arraycolsep=0pt
-                                      \begin{array}[t]{l}
-                                        % "red conditions" (in reference to blackboard picture)
-                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
-                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
-                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
-                                        \nonExec{cb_2} \text{ and } \\
-                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
-                                        % "yellow conditions"
-                                        \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
-                                        \noRetStkReg{\reg} \text{ and } \\
-                                        \noRetStkMs{\ms_\stk'} \text{ and } \\
-                                        \reg' = \reg
-                                        \begin{array}[t]{l}
-                                        \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
-                                        \update{\rretc}{\retptr}\\
-                                        \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
-                                        \end{array}
-\\                                        \text{ for $i \in \{1,2\}$ if $\isLinear{\Phi(r_i)}$, then $\reg''(r_i) = 0$ and} \\
-                                        \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
-                                        \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
-                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
-                                        \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
-                                      \end{array}\\
-                                      & \\
-      \tundefined & \arraycolsep=0pt
-                                      \begin{array}[t]{l}
-                                        % "red conditions" (reference to blackboard picture)
-                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
-                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
-                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
-                                        \nonExec{cb_2} \text{ and } \\
-                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
-                                        % "yellow conditions"
-                                        \text{not } \arraycolsep=0pt
-                                        \left(\begin{array}{l}
-                                          \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
-                                          \noRetStkReg{\reg} \text{ and } \\
-                                          \noRetStkMs{\ms_\stk'}
-                                        \end{array} \right)
-                                      \end{array}\\
-                                      & \\
-      \failed & \text{otherwise}
-    \end{cases}
-\end{align*}
-
-Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$.
-We cannot just use $\sjmp{}$ because it only takes one argument and we need to be able to jump to a code/data pair.
-To this end we introduce $\sxjmp{}{}$:
-\begin{align*}
-  \ssem{\sxjmp{r_1}{r_2}} = & 
-                              \begin{cases}
-                                \Phi'
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  [\reg.\pcreg \mapsto c_1] \\
-                                  [\reg.\rdata \mapsto c_2]
-                                \end{array} & 
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
-                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
-                                  \text{ for $i \in \{1,2\}$ if $\isLinear{c_i}$, then $\Phi'(r_i) = 0$} \\
-                                  \text{ otherwise $\Phi'(r_i) = \Phi(r_i)$ and} \\
-                                  \text{ $ \Phi(r) = \Phi'(r) \text{ for } r\in \tRegName \setminus \{r_1,r_2\}$ and} \\
-                                  \nonExec{c_2}
-                                \end{array}\\\\
-%
-                                (\mem,\reg',\stk',\ms_\stk') &
-                                \arraycolsep=0pt
-                                \begin{array}[t]{l}
-                                  \Phi(r_1) = \retptrc \text{ and } \\
-                                  \Phi(r_2) = \retptrd \text{ and } \\
-                                  \Phi = (\mem, \reg, \stk, \ms_\stk) \text{ and } \\
-                                  \stk = (\opc,\ms_\priv,\aaddr_\stk) :: \stk' \text{ and } \\
-                                  \reg \arraycolsep=0pt
-                                  \begin{array}[t]{l}
-                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\update{r_2}{0}\\
-                                  \end{array}
-                                \end{array}\\\\
-%
-                                \failed & \text{otherwise}
-                              \end{cases}
-\end{align*}
-\lau{30-08-2017: questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}
-
-Notice, $\rdata$ serves the same function as the IDC register in CHERI, i.e.\ this is where the data capability is put.
-In the interpretation of $\sxjmp{}{}$, we require $c_2$ to be non-executable because CHERI does it.
-We are, however, not certain whether this requirement is strictly necessary (maybe it is dangerous to $\sxjmp{r}{r}$ where $r$ points to the same capability?
-Maybe it can be use to unseal a code capability by first sealing another code capability with the seal, $\sxjmp{}{}$ with the newly sealed and the old sealed capability.
-Now both capabilities are unsealed).
-CHERI also checks whether $c_1$ is executable and in range, but we do this on every step of the operational semantics, so we do not repeat that check.
 

### 2017-09-21 fully-abstract-comp/technical-report.tex
typos
updated split and splice - now both do not work with empty capabilities - added some (informal) properties they should satisfy.

@@
 
-\textbf{split}\\
+\noindent\textbf{split and splice}\\
+We would like splice and split to have the following properties
+\begin{enumerate}
+\item No authority amplification - splitting or splicing capabilities should give you no more authority than you already had.
+\item Split should be dual to splice in the sence that a split on a capability followed by a splice of the two resulting capabilities should yield the same capability.
+\item Take the addresses governed by a linear capability to be a multiset.
+If this capability is split, then the union of the two multisets of addresses governed by the resulting capabilities should be the same as the first multiset.
+In other words, splice and split should not break linearity.
+\end{enumerate}
+Split cannot create ``empty capabilities'' (a capability that governs no segment of the memory, i.e.\ a capability where the base address is greater than the end address).
+We partly do not allow this out of convenience as it makes the implementation of call simpler.
+We do not need empty capabilities as they have no semantic value in the sense that they allow you to do no more than a piece of data.
 \begin{align*}
@@
                                    \quad\text{ and in either case } n \in \nats \\
-                                   \baddr \leq n \tand n \leq \eaddr \tand \\
-                                   c_1 = ((\perm,\lin),\baddr,n,\baddr) \tand \\
-                                   c_2 = ((\perm,\lin),n+1,\eaddr,n+1) \tand \\
+                                   \baddr \leq n \tand n < \eaddr \tand \\
+                                   c_1 = ((\perm,\lin),\baddr,n,\aaddr) \tand \\
+                                   c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr) \tand \\
                                    w = \linCons{\Phi(r_3)}
@@
                                    \Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                   \perm \neq \enter \\
                                    \Phi(r_4) = n \tor r_4 = n \\
                                    \quad\text{ and in either case } n \in \nats \\
-                                   \baddr \leq n \tand n \leq \eaddr \tand \\
-                                   c_1 = \stkptr{\perm,\baddr,n,\baddr} \tand \\
-                                   c_2 = \stkptr{\perm,n+1,\eaddr,n+1} 
+                                   \baddr \leq n \tand n < \eaddr \tand \\
+                                   c_1 = \stkptr{\perm,\baddr,n,\aaddr} \tand \\
+                                   c_2 = \stkptr{\perm,n+1,\eaddr,\aaddr} 
                                  \end{array} \right.\\
@@
 
-\noindent\textbf{splice}\\
 \lau{19-09-2017: Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
@@
                                   \Phi(r_2) = ((\perm,\lin),\baddr_2,\eaddr_2,\aaddr_2) \tand \\
-                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\aaddr_3) \tand \\
+                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\_) \tand \\
                                   \perm \neq \enter \tand \\
@@
                                                  \begin{array}[t]{l}
-                                                   \Phi(r_1) = \sealed(c_1,\sigma) \tand \\
-                                                   \Phi(r_2) = \sealed(c_2,\sigma) \tand \\
+                                                   \Phi(r_1) = \sealed{c_1,\sigma} \tand \\
+                                                   \Phi(r_2) = \sealed{c_2,\sigma} \tand \\
                                                    \nonExec{c_2} \tand\\

### 2017-09-21 fully-abstract-comp/technical-report.tex
Updated sealing-tokens to work more like capabilities.
Added a new setup for linking/program layout. Removed the old section

@@
 }
+
+% Rotate entire figures
+\usepackage{rotating}
 
@@
     &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
-    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf}\\
+    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
@@
       \cb = ((\_,\linear),\_,\_,\_) \tor\\
-      \cb = \seal{\linear,\_,\_} \tor\\
+      \cb = \seal{\linear,\_,\_,\_} \tor\\
       \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
@@
                                        \Phi(r_2) = ((\_,\_),\_,\_,\aaddr) \\
+                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\_,\aaddr} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\aaddr} } \\
@@
                                        \Phi(r_2) = ((\_,\_),\baddr,\_,\_) \\
+                                       \quad \tor \Phi(r_2) = \seal{\_,\baddr,\_,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\baddr,\_,\_} } \\
@@
                                        \Phi(r_2) = ((\_,\_),\_,\eaddr,\_) \\
+                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\eaddr,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\eaddr,\_} } \\
@@
                                        \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
+                                       \quad \tor \Phi(r_2) = \seal{\lin,\_,\_,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} \tand l = \linear}\\
@@
                                     \\
+                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                      \quad\text{ and in either case $n \in \ints$ and} \\
+                                      \Phi(r_1) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
+                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
+                                    \end{array}
+                                    \\
                                     \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
                                       \perm \neq \enter \tand \\
@@
                                       c = ((\perm,\lin),\baddr,\eaddr,\baddr)
+                                    \end{array} \\
+                                  \updPcAddr{}\left(
+                                    \arraycolsep=0pt
+                                    \begin{array}{rl}
+                                    \Phi&\updReg{r_2}{w_2}\\
+                                        &\updReg{r_1}{c}
+                                    \end{array} \right)
+&
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
+                                      w_2 = \linCons{\Phi(r_2)}\\
+                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
                                     \end{array} \\
@@
                                       \Phi(r_2) \in \SealableCaps \tand \\
-                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr} \tand \\
-                                      \sigma_\baddr \leq \sigma_\eaddr \tand \\ 
+                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
+                                      \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
                                       \vsc = \sealed{c,\sigma}
@@
                                  \begin{array}[t]{l}
-                                   \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
+                                   \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
                                    \Phi(r_4) = n \tor r_4 = n \\
@@
                                    \sigma_\baddr \leq n \tand n < \sigma_\eaddr \tand \\
-                                   c_1 = \seal{\lin,\sigma_\baddr,n} \tand \\
-                                   c_2 = \seal{\lin,n+1,\sigma_\eaddr} \tand \\
+                                   c_1 = \seal{\lin,\sigma_\baddr,n,\sigma} \tand \\
+                                   c_2 = \seal{\lin,n+1,\sigma_\eaddr,\sigma} \tand \\
                                    w = \linCons{\Phi(r_3)}
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2}} \tand \\
-                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3}} \tand \\
+                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2},\sigma} \tand \\
+                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3},\_} \tand \\
                                   \sigma_{\eaddr,2}+1 = \sigma_{\baddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand\\
                                   \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
-                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3}} \tand \\
+                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3},\sigma} \tand \\
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
@@
 
-\subsection{Target program linking and layout}
-\lau{13-09-2017: TODO take another look and linking and consider what the setup must be now that we don't have local capabilities.}
-For a $\scall{}{}$ to be secure after compilation, we need to use the seal mechanism.
-In order for a program to seal a capability, it must have a seal available.
-We propose that a compiled program must have the layout sketched in Figure~\ref{fig:trg-prog-link}.
-The first address of a program contains a seal (unique to this program), the second address contains a capability for a linking table, and the following addresses contain the instructions of the program.
-The capability for the program points to the first address of the program and it has permission no greater then global read-execute (notice that it does not have write).
-
-The capability for the linking table is a global read-write capability, which means that it does not grant access to alter the table (which means that the table could potentially be shared between multiple parties).
-The linking table contains pairs of sealed code and data capabilities.
-
-The program should also have access to a persistent memory which it has in the register $\rdata$ where it has a global read-write capability (or something less) (notice that this capabilty should not have execute permission).
-Finally the program also has access to local memory in the form of a stack.
-The capability for the stack resides in $\rstk$ and the capability should have local read-write-local permission (or less)
-\lau{30-08-2017: The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}
-
-\lau{30-08-2017: should the capability for the ``persistent'' memory reside along side the code?}
-
-Notice that the linking part is not strictly necessary for the compilation, but we would like to be able to link our known programs with other programs.
-
+\subsection{Program layout and linking}
+A target-level program needs a number of things available in order to be able to execute securely. In the following, we list what it needs followed by a description of how they are made available.
+
+A program needs access to the following:
+\begin{description}
+\item[Code] The instructions of a program should of course be available so they can be executed.
+\item[Program seals] Each call/return point of the program needs a unique seal in order to enforce well-bracketedness. This seal needs to be unique on the entire machine.
+\item[Stack] A program stores the value of local variables on the stack. 
+\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory seperate from the stack.
+\item[Linking table] A progam may need to call other programs. Other programs are made available in a linking table.
+\end{description}
+
+In the following, we describe the setup in detail. Figure~\ref{fig:trg-prog-link} illustrates the setup.
+
+The \emph{code}, \emph{program seals}, and \emph{linking table} are available through a normal read-execute capability in the $\pc$ register.
+The program seals are found on the first address governed by this capability.
+The second address governe by the capability contains a normal read capability that governs the linking table and points to the first address.
+The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
+
+The \emph{linking table} contains pairs of sealed code and data capabilities. So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
+
+The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.} The seals are globally unique in the sense that no other program has access to this range of seals.
+
+The \emph{stack} is available through a capability in the register $r_\stk$.
+The stack pointer is a linear read write capability.
+There are three ways of receiving a stack: (1) the first program executing has a stack pointer in the $r_\stk$ register, (2) call backs and programs can expect to be called with a stack pointer as an argument in the $r_\stk$ register, and (3) when returning from a call, part of the stack is unsealed as the data part of a return pointer the other part is returned from the callee (the callee is allowed to keep part of the stack, but the callee must at least return a non-empty part of the stack adjacent to the part of the stack that was unsealed by returning).
+
+The \emph{peristent data storage} is expected to be accisble through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
+There are two ways for a program to get this data capability: (1) the data part of the code/data part should be the sealed data capability.
+(2) During a call, the data capability should be stored on the stack, so it can be retrieved afterwards.
 \begin{figure}
   \centering
-  \includegraphics[width=\textwidth]{img/prog_layout.jpg}
-  \caption{Target program linking and layout (there is a typo in the figure: the pc capability should be global rather than local).}
+  \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
+  \caption{Linking and program layout.}
   \label{fig:trg-prog-link}

### 2017-09-22 fully-abstract-comp/technical-report.tex
Update to xjmp to allow callee to return stack of any size. Also typos.

@@
                                       \withinBounds{\Phi(r_2)} \tand \\
-                                      \in \dom(\Phi.\ms_\stk) \tand \\
+                                      \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
                                       w = \Phi.\ms_\stk(a) \tand \\
@@
 \noindent\textbf{xjmp}\\
-\lau{18-09-2017: Work in progress - remember to delete this when sem. done.}
+\lau{22-09-2017: Can we allow the callee to return any stack pointer?
+At the moment, it is not possible to create an empty stack pointer which may be a saving grace.
+If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
 \begin{align*}
@@
                                          c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
-                                         \dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk]\tand \\
-                                         \dom(\ms_{\stk,\priv} = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
+                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
@@
                                       \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
-                                      \vsc = \sealed{c,\sigma}
+                                      \vsc = \sealed{\Phi(r_2),\sigma}
                                     \end{array}
@@
                                                  \begin{array}[t]{l}
-                                                   \Phi(r_1) = \sealed{c_1,\sigma} \tand \\
-                                                   \Phi(r_2) = \sealed{c_2,\sigma} \tand \\
+                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
+                                                   \Phi(r_2) = \sealed{c_2,\sigma_1} \tand \\
+                                                   \sigma_1 = \sigma_2 \tand \\
                                                    \nonExec{c_2} \tand\\

### 2017-09-25 fully-abstract-comp/technical-report.tex
added missing linCons to cseal instruction
updated call semantics to not require all of the stack
Added small section about why callee must return all of the stack.

@@
                                          \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
+                                         \dom(\ms_{\stk,\priv} \uplus \ms_\stk) = [\baddr_{\stk,\priv},\eaddr_\stk] \tand \\
                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
@@
                                   \begin{cases}
-                                    \Phi\updReg{r_1}{\vsc} &
+                                    \arraycolsep=0pt
+                                    \begin{array}{rl}
+                                      \Phi&\updReg{r_1}{\vsc}\\
+                                           \updReg{r_2}{w_2}
+                                    \end{array}
+&
                                     \arraycolsep=0pt
@@
                                       \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
-                                      \vsc = \sealed{\Phi(r_2),\sigma}
+                                      \vsc = \sealed{\Phi(r_2),\sigma} \tand\\
+                                      w_2 = \linCons{\Phi(r_2)}                                     
                                     \end{array}
@@
 
+\subsection{Calling convention design decisions}
+\subsubsection{Returning the full stack}
+When a callee return from a call, they must return all of the stack they were passed.
+If we omit this requirement, then we cannot guarantee well-bracketedness.
+The following is an example of circumventing well-bracketedness by keeping part of the stack:
+
+\begin{itemize}
+\item An adversary calls our trusted code with a call-back. 
+\item Our code uses part of the stack and calls the callback with the rest of the stack.
+\item The adversary splits that stack in two and saves the part of the stack adjacent to our stack in some persistent memory.
+The adversary calls us anew with a callback and the part of the stack they did not save.
+\item We use part of the stack we receive and call the callback with the rest of the stack.
+The adversary can now use the saved stack to return from the first call to the callback breaking well-bracketedness.
+The reason this is possible is because we do not check the size of the stack.
+\end{itemize}
+Figure~\ref{fig:ret-full-stk} illustrates the above example.
+In the figure, $f$ is the function to ``is'' and $g$ is the callback passed to ``us''.
+In the end, the part of the stack marked \emph{kept by adv} can be used to return from the first call-back (the call with the *).
+The stack that is used to return lines up with $\var{Us}_\priv$, so the return is successful.
+\begin{figure}
+  \centering
+  \includegraphics[angle=90,width=\textwidth]{img/ret-full-stk.pdf}
+  \caption{Illustration of the stack in the example that illustrates why the entire stack must be returned.}
+  \label{fig:ret-full-stk}
+\end{figure}
+
 \end{document}

### 2017-09-25 fully-abstract-comp/technical-report.tex
Updated call and xjmp semantics to be more 'relaxed'
updated call with a condition related to the seal

@@
 \newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
+\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
 \newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
@@
                                          \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
+                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
+                                         \ms_{\stk,\priv}(\baddr_\stk-1) = \eaddr_\stk \tand \\
                                          c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
-                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
-                                         \dom(\ms_{\stk,\priv} \uplus \ms_\stk) = [\baddr_{\stk,\priv},\eaddr_\stk] \tand \\
                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
@@
 \lau{19-09-2017: In which cases do we have undefined here?}
+This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack.
+We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}
 \begin{align*}
@@
                                                    \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
-                                                   \Phi(r_2) = \sealed{c_2,\sigma_1} \tand \\
+                                                   \Phi(r_2) = \sealed{c_2,\sigma_2} \tand \\
                                                    \sigma_1 = \sigma_2 \tand \\
                                                    \nonExec{c_2} \tand\\
+                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
                                                    \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
-                                                   \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
-                                                   \baddr_\stk \leq \aaddr_\stk \leq \eaddr_\stk \tand \\
-                                                   \dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk] \tand\\
+                                                   \baddr_\stk \leq \aaddr_\stk < \eaddr_\stk \tand \\
                                                    \ms_{\stk,\priv} = \ms_\stk |_{[\baddr_\stk,\aaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
                                                    \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand\\
+                                                   c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
                                                    \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                                    \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
-                                                   c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
-                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk)\tand\\
+                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk) :: \stk \tand\\
                                                    \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
+                                                   \mem(\baddr) = \seal{\_,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
+                                                   \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand
+                                                   \sigma_? = \sealAss{a} \tand\\
                                                    w_1 = \linCons{\Phi(r_1)} \tand \\

### 2017-09-25 fully-abstract-comp/technical-report.tex
Add a remark and a small fix.

--- 2017-09-22 fully-abstract-comp/technical-report.tex
@@
 }
-
-% Rotate entire figures
-\usepackage{rotating}
 
@@
 % Four arguments
-%\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
+\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
 \newcommand{\ssplit}[4]{\fourinstr{\src{split}}{#1}{#2}{#3}{#4}}
@@
 % Four arguments
-%\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
+\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
 \newcommand{\tsplit}[4]{\fourinstr{\trg{split}}{#1}{#2}{#3}{#4}}
@@
     &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
-    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
+    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf}\\
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
@@
     &\MemSeg & \defeq & \Addr \parfun \Word \\
-    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
+    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
     \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tsubseg{\trg{r}}{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
   & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
@@
       \cb = ((\_,\linear),\_,\_,\_) \tor\\
-      \cb = \seal{\linear,\_,\_,\_} \tor\\
+      \cb = \seal{\linear,\_,\_} \tor\\
       \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
@@
 }
+\dominique{20-9-2017: I also think gettype for source-specific words should produce the same result than gettype for their implementation.}
 \begin{align*}
@@
                                        \Phi(r_2) = ((\_,\_),\_,\_,\aaddr) \\
-                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\_,\aaddr} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\aaddr} } \\
@@
                                        \Phi(r_2) = ((\_,\_),\baddr,\_,\_) \\
-                                       \quad \tor \Phi(r_2) = \seal{\_,\baddr,\_,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\baddr,\_,\_} } \\
@@
                                        \Phi(r_2) = ((\_,\_),\_,\eaddr,\_) \\
-                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\eaddr,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\eaddr,\_} } \\
@@
                                        \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                       \quad \tor \Phi(r_2) = \seal{\lin,\_,\_,\_} \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} \tand l = \linear}\\
+                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} } \tand l = \linear\\
                                      \end{array} \\
@@
                               \begin{cases}
-                                \updPcAddr{\Phi\updReg{r_1}{r_2}} & r_2 \in \ints \\
                                 \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
@@
                                       \begin{array}{rl}
-                                        \Phi&\update{\mem_\stk.a}{w}\\
+                                        \Phi&\update{\mem.a}{w}\\
                                             &\updReg{r_2}{w_2}
@@
                                       \withinBounds{\Phi(r_2)} \tand \\
-                                      \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
+                                      \in \dom(\Phi.\ms_\stk) \tand \\
                                       w = \Phi.\ms_\stk(a) \tand \\
@@
                                       c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
-                                    \end{array}
-                                    \\
-                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
-                                    \arraycolsep=0pt
-                                    \begin{array}[t]{l}
-                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
-                                      \quad\text{ and in either case $n \in \ints$ and} \\
-                                      \Phi(r_1) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
-                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
                                     \end{array}
@@
 
+\noindent\textbf{subseg}\\
+\lau{19-09-2017: TODO: I am not sure the way the end points are handled makes entirely sense. Have a look at it (maybe make a function that computes the end point?)}
+\begin{align*}
+  \sem{\tsubseg{r_1}{r_2}{r_3}{r_4}}(\Phi) = &
+                                               \begin{cases}
+                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \text{for $i \in \{3, 4\}$ } \\
+                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
+                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
+                                                   \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                                   \baddr \leq n_3 \tand \\
+                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
+                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
+                                                   \perm \neq \enter \tand\\
+                                                   c = ((\perm,\lin),n_1,\eaddr',\aaddr)\\
+                                                   w_2 = \linCons{\Phi(r_2)}
+                                                 \end{array}
+\\
+                                                 \Phi\updReg{r_2}{w_2}\updReg{r_1}{c} &
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \text{for $i \in \{3, 4\}$ } \\
+                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
+                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
+                                                   \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
+                                                   \baddr \leq n_3 \tand \\
+                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand \\
+                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
+                                                   \perm \neq \enter \tand\\
+                                                   c = \seal{\lin,n_1,\eaddr'}\\
+                                                   w_2 = \linCons{\Phi(r_2)}
+                                                 \end{array}
+\\
+                                                 \src{\Phi\updReg{r_2}{0}\updReg{r_1}{c}} &
+                                                 \sourcecolor\left.
+                                                 \arraycolsep=0pt
+                                                 \begin{array}[t]{l}
+                                                   \text{for $i \in \{3, 4\}$ } \\
+                                                   \quad \text{either $n_i = r_i$ or $n_i = \Phi.\reg(r_i)$}\\
+                                                   \quad \text{and in either case $n_i \in \ints$ and} \\
+                                                   \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                                   \baddr \leq n_3 \tand \\
+                                                   (n_2 \leq \eaddr \tand n_2 \in \nats) \Rightarrow \eaddr' = n_2 \tand\\
+                                                   (n_2 = -42 \tand \eaddr = \infty) \Rightarrow \eaddr' = \infty \tand\\
+                                                   \perm \neq \enter \tand\\
+                                                   c = \stkptr{\perm,n_1,\eaddr',\aaddr}\\
+                                                 \end{array}\right.\\
+                                                 \failed & \totherwise 
+                                               \end{cases}
+\end{align*}
+
 \noindent\textbf{lt}\\
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
+                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                       \perm \neq \enter \tand \\
@@
                                     \end{array} \\
-                                  \updPcAddr{}\left(
-                                    \arraycolsep=0pt
-                                    \begin{array}{rl}
-                                    \Phi&\updReg{r_2}{w_2}\\
-                                        &\updReg{r_1}{c}
-                                    \end{array} \right)
-&
-                                    \arraycolsep=0pt
-                                    \begin{array}[t]{l}
-                                      \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
-                                      w_2 = \linCons{\Phi(r_2)}\\
-                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
-                                    \end{array} \\
                                     \failed & \totherwise
@@
 \noindent\textbf{xjmp}\\
-\lau{22-09-2017: Can we allow the callee to return any stack pointer?
-At the moment, it is not possible to create an empty stack pointer which may be a saving grace.
-If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
+\lau{18-09-2017: Work in progress - remember to delete this when sem. done.}
 \begin{align*}
@@
                                          c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
-                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
+                                         \dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk]\tand \\
+                                         \dom(\ms_{\stk,\priv} = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
@@
                                       \Phi(r_2) \in \SealableCaps \tand \\
-                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
-                                      \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
-                                      \vsc = \sealed{\Phi(r_2),\sigma}
+                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr} \tand \\
+                                      \sigma_\baddr \leq \sigma_\eaddr \tand \\ 
+                                      \vsc = \sealed{c,\sigma}
                                     \end{array}
@@
 
-\noindent\textbf{split and splice}\\
-We would like splice and split to have the following properties
-\begin{enumerate}
-\item No authority amplification - splitting or splicing capabilities should give you no more authority than you already had.
-\item Split should be dual to splice in the sence that a split on a capability followed by a splice of the two resulting capabilities should yield the same capability.
-\item Take the addresses governed by a linear capability to be a multiset.
-If this capability is split, then the union of the two multisets of addresses governed by the resulting capabilities should be the same as the first multiset.
-In other words, splice and split should not break linearity.
-\end{enumerate}
-Split cannot create ``empty capabilities'' (a capability that governs no segment of the memory, i.e.\ a capability where the base address is greater than the end address).
-We partly do not allow this out of convenience as it makes the implementation of call simpler.
-We do not need empty capabilities as they have no semantic value in the sense that they allow you to do no more than a piece of data.
+\textbf{split}\\
 \begin{align*}
@@
                                    \quad\text{ and in either case } n \in \nats \\
-                                   \baddr \leq n \tand n < \eaddr \tand \\
-                                   c_1 = ((\perm,\lin),\baddr,n,\aaddr) \tand \\
-                                   c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr) \tand \\
+                                   \baddr \leq n \tand n \leq \eaddr \tand \\
+                                   c_1 = ((\perm,\lin),\baddr,n,\baddr) \tand \\
+                                   c_2 = ((\perm,\lin),n+1,\eaddr,n+1) \tand \\
                                    w = \linCons{\Phi(r_3)}
@@
                                  \begin{array}[t]{l}
-                                   \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
+                                   \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr} \tand \\
                                    \Phi(r_4) = n \tor r_4 = n \\
@@
                                    \sigma_\baddr \leq n \tand n < \sigma_\eaddr \tand \\
-                                   c_1 = \seal{\lin,\sigma_\baddr,n,\sigma} \tand \\
-                                   c_2 = \seal{\lin,n+1,\sigma_\eaddr,\sigma} \tand \\
+                                   c_1 = \seal{\lin,\sigma_\baddr,n} \tand \\
+                                   c_2 = \seal{\lin,n+1,\sigma_\eaddr} \tand \\
                                    w = \linCons{\Phi(r_3)}
@@
                                    \Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                   \perm \neq \enter \\
                                    \Phi(r_4) = n \tor r_4 = n \\
                                    \quad\text{ and in either case } n \in \nats \\
-                                   \baddr \leq n \tand n < \eaddr \tand \\
-                                   c_1 = \stkptr{\perm,\baddr,n,\aaddr} \tand \\
-                                   c_2 = \stkptr{\perm,n+1,\eaddr,\aaddr} 
+                                   \baddr \leq n \tand n \leq \eaddr \tand \\
+                                   c_1 = \stkptr{\perm,\baddr,n,\baddr} \tand \\
+                                   c_2 = \stkptr{\perm,n+1,\eaddr,n+1} 
                                  \end{array} \right.\\
@@
 
+\noindent\textbf{splice}\\
 \lau{19-09-2017: Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
@@
                                   \Phi(r_2) = ((\perm,\lin),\baddr_2,\eaddr_2,\aaddr_2) \tand \\
-                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\_) \tand \\
+                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\aaddr_3) \tand \\
                                   \perm \neq \enter \tand \\
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2},\sigma} \tand \\
-                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3},\_} \tand \\
+                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2}} \tand \\
+                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3}} \tand \\
                                   \sigma_{\eaddr,2}+1 = \sigma_{\baddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand\\
                                   \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
-                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3},\sigma} \tand \\
+                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3}} \tand \\
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
@@
                                   \perm \neq \enter \tand \\
-                                  \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
+                                  \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr2 \tand \baddr_3 \leq \eaddr_3 \tand \\
                                   c = \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_2} \\
@@
                                                  \begin{array}[t]{l}
-                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
-                                                   \Phi(r_2) = \sealed{c_2,\sigma_1} \tand \\
-                                                   \sigma_1 = \sigma_2 \tand \\
+                                                   \Phi(r_1) = \sealed(c_1,\sigma) \tand \\
+                                                   \Phi(r_2) = \sealed(c_2,\sigma) \tand \\
                                                    \nonExec{c_2} \tand\\
@@
 
-\subsection{Program layout and linking}
-A target-level program needs a number of things available in order to be able to execute securely. In the following, we list what it needs followed by a description of how they are made available.
-
-A program needs access to the following:
-\begin{description}
-\item[Code] The instructions of a program should of course be available so they can be executed.
-\item[Program seals] Each call/return point of the program needs a unique seal in order to enforce well-bracketedness. This seal needs to be unique on the entire machine.
-\item[Stack] A program stores the value of local variables on the stack. 
-\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory seperate from the stack.
-\item[Linking table] A progam may need to call other programs. Other programs are made available in a linking table.
-\end{description}
-
-In the following, we describe the setup in detail. Figure~\ref{fig:trg-prog-link} illustrates the setup.
-
-The \emph{code}, \emph{program seals}, and \emph{linking table} are available through a normal read-execute capability in the $\pc$ register.
-The program seals are found on the first address governed by this capability.
-The second address governe by the capability contains a normal read capability that governs the linking table and points to the first address.
-The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
-
-The \emph{linking table} contains pairs of sealed code and data capabilities. So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
-
-The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.} The seals are globally unique in the sense that no other program has access to this range of seals.
-
-The \emph{stack} is available through a capability in the register $r_\stk$.
-The stack pointer is a linear read write capability.
-There are three ways of receiving a stack: (1) the first program executing has a stack pointer in the $r_\stk$ register, (2) call backs and programs can expect to be called with a stack pointer as an argument in the $r_\stk$ register, and (3) when returning from a call, part of the stack is unsealed as the data part of a return pointer the other part is returned from the callee (the callee is allowed to keep part of the stack, but the callee must at least return a non-empty part of the stack adjacent to the part of the stack that was unsealed by returning).
-
-The \emph{peristent data storage} is expected to be accisble through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
-There are two ways for a program to get this data capability: (1) the data part of the code/data part should be the sealed data capability.
-(2) During a call, the data capability should be stored on the stack, so it can be retrieved afterwards.
+\subsubsection{Source Language}
+\label{sec:source-language}
+\lau{14-09-2017: This section will be obsolete when the unifyed definitions are done.}
+\lau{23-03-2017: $\sjmp{}$ and $\scall{}{}$ needs an update. Update: they have been updated (It was $\sxjmp{}{}$ that needed an update.}
+
+\textbf{Call}\\
+Call has three cases, a successful call, undefined and failed.
+For a call to be successful, it needs two sealed capabilities - a data and a code capability.
+The code capability must be non-executable and the stack register must contain a stack pointer.
+\lau{30-08-2017: Add more description here.}
+
+For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee,
+\dominique{7-9-2017: We declare the call to be undefined if the caller is doing something ``stupid'' that would break security but which we do not want to check for.
+This includes: leaving a copy of the stack pointer or the old return pointer in a place accessible to the callee, invoking local callbacks (see discussion in POPL2018 submission of why this is bad).
+The condition that $\cb_2 \neq \stkptr{\_,\_,\_,\_}$ is weird because this would not be the caller's fault, but the callee's.
+Further thought needed regarding correct blame assignment for this case.}
+
+If we are not in one of the two cases above, then the execution fails.
+\begin{align*}
+  \ssem{\scall{r_1}{r_2}} = & 
+    \begin{cases}
+      (\mem,\reg',\stk',\ms_\stk') & \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        % "red conditions" (in reference to blackboard picture)
+                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
+                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
+                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
+                                        \nonExec{cb_2} \text{ and } \\
+                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
+                                        % "yellow conditions"
+                                        \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
+                                        \noRetStkReg{\reg} \text{ and } \\
+                                        \noRetStkMs{\ms_\stk'} \text{ and } \\
+                                        \reg' = \reg
+                                        \begin{array}[t]{l}
+                                        \update{\pcreg}{\cb_1}\update{\rdata}{\cb_2}\\
+                                        \update{\rretc}{\retptr}\\
+                                        \update{\rstk}{\stkptr{\perm,[\aaddr,\eaddr],\aaddr}} \text{ and }
+                                        \end{array}
+\\                                        \text{ for $i \in \{1,2\}$ if $\isLinear{\Phi(r_i)}$, then $\reg''(r_i) = 0$ and} \\
+                                        \opc = \Phi(\pcreg)\update{a}{\Phi(\pcreg).a+\calllen} \text{ and } \\
+                                        \ms_\priv = \ms_\stk \setminus [\aaddr,\eaddr] \text{ and } \\
+                                        \stk' = (\opc,\ms_\priv,\perm,\stdrng,\aaddr) :: \stk \text{ and } \\
+                                        \ms_\stk' = \ms_\stk |_{[\baddr,\eaddr]}
+                                      \end{array}\\
+                                      & \\
+      \tundefined & \arraycolsep=0pt
+                                      \begin{array}[t]{l}
+                                        % "red conditions" (reference to blackboard picture)
+                                        \Phi = (\mem,\reg,\stk,\ms_\stk) \text{ and } \\
+                                        \Phi(r_1) = \sealed{\cb_1,\sigma} \text{ and } \\
+                                        \Phi(r_2) = \sealed{\cb_2,\sigma} \text{ and } \\
+                                        \nonExec{cb_2} \text{ and } \\
+                                        \Phi(\rstk) = \stkptr{\perm, \stdrng, \aaddr} \text{ and } \\
+                                        % "yellow conditions"
+                                        \text{not } \arraycolsep=0pt
+                                        \left(\begin{array}{l}
+                                          \cb_2 \neq \stkptr{\_,\_,\_,\_} \text{ and } \\
+                                          \noRetStkReg{\reg} \text{ and } \\
+                                          \noRetStkMs{\ms_\stk'}
+                                        \end{array} \right)
+                                      \end{array}\\
+                                      & \\
+      \failed & \text{otherwise}
+    \end{cases}
+\end{align*}
+
+Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$.
+We cannot just use $\sjmp{}$ because it only takes one argument and we need to be able to jump to a code/data pair.
+To this end we introduce $\sxjmp{}{}$:
+\begin{align*}
+  \ssem{\sxjmp{r_1}{r_2}} = & 
+                              \begin{cases}
+                                \Phi'
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  [\reg.\pcreg \mapsto c_1] \\
+                                  [\reg.\rdata \mapsto c_2]
+                                \end{array} & 
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
+                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
+                                  \text{ for $i \in \{1,2\}$ if $\isLinear{c_i}$, then $\Phi'(r_i) = 0$} \\
+                                  \text{ otherwise $\Phi'(r_i) = \Phi(r_i)$ and} \\
+                                  \text{ $ \Phi(r) = \Phi'(r) \text{ for } r\in \tRegName \setminus \{r_1,r_2\}$ and} \\
+                                  \nonExec{c_2}
+                                \end{array}\\\\
+%
+                                (\mem,\reg',\stk',\ms_\stk') &
+                                \arraycolsep=0pt
+                                \begin{array}[t]{l}
+                                  \Phi(r_1) = \retptrc \text{ and } \\
+                                  \Phi(r_2) = \retptrd \text{ and } \\
+                                  \Phi = (\mem, \reg, \stk, \ms_\stk) \text{ and } \\
+                                  \stk = (\opc,\ms_\priv,\aaddr_\stk) :: \stk' \text{ and } \\
+                                  \reg \arraycolsep=0pt
+                                  \begin{array}[t]{l}
+                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\update{r_2}{0}\\
+                                  \end{array}
+                                \end{array}\\\\
+%
+                                \failed & \text{otherwise}
+                              \end{cases}
+\end{align*}
+\lau{30-08-2017: questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}
+
+Notice, $\rdata$ serves the same function as the IDC register in CHERI, i.e.\ this is where the data capability is put.
+In the interpretation of $\sxjmp{}{}$, we require $c_2$ to be non-executable because CHERI does it.
+We are, however, not certain whether this requirement is strictly necessary (maybe it is dangerous to $\sxjmp{r}{r}$ where $r$ points to the same capability?
+Maybe it can be use to unseal a code capability by first sealing another code capability with the seal, $\sxjmp{}{}$ with the newly sealed and the old sealed capability.
+Now both capabilities are unsealed).
+CHERI also checks whether $c_1$ is executable and in range, but we do this on every step of the operational semantics, so we do not repeat that check.
+
+\subsection{Target program linking and layout}
+\lau{13-09-2017: TODO take another look and linking and consider what the setup must be now that we don't have local capabilities.}
+For a $\scall{}{}$ to be secure after compilation, we need to use the seal mechanism.
+In order for a program to seal a capability, it must have a seal available.
+We propose that a compiled program must have the layout sketched in Figure~\ref{fig:trg-prog-link}.
+The first address of a program contains a seal (unique to this program), the second address contains a capability for a linking table, and the following addresses contain the instructions of the program.
+The capability for the program points to the first address of the program and it has permission no greater then global read-execute (notice that it does not have write).
+
+The capability for the linking table is a global read-write capability, which means that it does not grant access to alter the table (which means that the table could potentially be shared between multiple parties).
+The linking table contains pairs of sealed code and data capabilities.
+
+The program should also have access to a persistent memory which it has in the register $\rdata$ where it has a global read-write capability (or something less) (notice that this capabilty should not have execute permission).
+Finally the program also has access to local memory in the form of a stack.
+The capability for the stack resides in $\rstk$ and the capability should have local read-write-local permission (or less)
+\lau{30-08-2017: The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}
+
+\lau{30-08-2017: should the capability for the ``persistent'' memory reside along side the code?}
+
+Notice that the linking part is not strictly necessary for the compilation, but we would like to be able to link our known programs with other programs.
+
 \begin{figure}
   \centering
-  \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
-  \caption{Linking and program layout.}
+  \includegraphics[width=\textwidth]{img/prog_layout.jpg}
+  \caption{Target program linking and layout (there is a typo in the figure: the pc capability should be global rather than local).}
   \label{fig:trg-prog-link}

### 2017-09-25 fully-abstract-comp/technical-report.tex
Remove some old comments.

--- 2017-09-25 fully-abstract-comp/technical-report.tex
@@
 
-\dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
-\lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost).
+% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
+% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost).
 I do, however, not think that it is strictly necessary here to make things work.}
-\lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
+% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
 Source language domains:

### 2017-09-25 fully-abstract-comp/technical-report.tex
added a subsection about allocation along with the example

--- 2017-09-25 fully-abstract-comp/technical-report.tex
@@
                                                    \ms_{\stk,\priv} = \ms_\stk |_{[\baddr_\stk,\aaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
-                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand\\
+                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand \\
                                                    c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
@@
                                                    \mem(\baddr) = \seal{\_,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
-                                                   \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand
+                                                   \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
                                                    \sigma_? = \sealAss{a} \tand\\
@@
 
+\subsubsection{Restriction on stack allocation}
+We need to somehow make sure that it is always the same stack that is used.
+If we don't, then an adversary can simply split the stack in two, use one part for one call and the other for another call.
+At this point, they can return to either of the two calls - in other words, well-bracketedness is not enforced.
+
+\begin{itemize}
+  \item An adversary starts the execution. They split the stack in two and call us with one part (say the top part) along with a callback.
+  \item We use part of the stack and call the adversary with the rest of the stack.
+  \item The adversary calls us again this time using the other part of the stack (here the bottom part of the stack).
+  \item Again, we use part of the stack and call the adversary with the rest of this part of the stack.
+\end{itemize}
+At this point, the adversary can return from either of the two calls. Swapping around the order in which the adversary uses the two parts of the stack changes nothing.
+
+The example is illustrated in Figure~\ref{fig:stk-alloc}.
+
+One way to solve this is to make it publicly known what the end address of the stack is and always check that we receive a part of the stack that goes all the way to the end.
+In the above example (using the stack parts in parenthesis) this would cause the sequence of calls to fail the second time we are called, because we do not receive a stack that goes all the way to the top of the stack.
+
+\begin{figure}
+  \centering
+  \includegraphics[angle=270,trim={0 0 5cm 0},width=\textwidth]{img/stk-alloc.pdf}
+  \caption{Illustration of the potential stack allocation issue.}
+  \label{fig:stk-alloc}
+\end{figure}
+
 \end{document}

### 2017-09-26 fully-abstract-comp/technical-report.tex
fixed store bug ?
Add a comment and fix a typo.

@@
 (I think it may have been from when we had call and return instructions?)}
+\dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
 There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
@@
 \subsubsection{Notes}
-Genrally:
+Generally:
 \begin{itemize}
@@
                                       \begin{array}{rl}
-                                        \Phi&\update{\mem.a}{w}\\
-                                            &\updReg{r_2}{w_2}
+                                        \Phi&\updReg{r_2}{w_2}\\
+                                            &\update{\ms_\stk.a}{w}
                                       \end{array}
@@
                                         \withinBounds{\Phi(r_1)} \tand \\
-                                        a \in \dom(\Phi.\ms_\stk) \tand \\
+                                        \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
                                         w = \Phi(r_2) \\

### 2017-09-27 fully-abstract-comp/technical-report.tex
More remarks and typo's.

@@
 \lau{13-09-2017: For now I have added linearity to the seals.}
+\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
+\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
 The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
@@
 
+\dominique{I think we need to do something about tail calls in the source language?}
+
 Target language:
@@
 We assume function to encode and decode permissions as well as a function to encode and decode linearity.
-The functions are used implicitely when a permission or linearity is used in a place where they need to be a word.
+The functions are used implicitly when a permission or linearity is used in a place where they need to be a word.
 \lau{14-09-2017: Write more about this (types and assumptions)}

### 2017-09-27 fully-abstract-comp/technical-report.tex
Fix typos, add a comment.

@@
 We assume functions $\decPerm{}$ and $\encPerm{}$.
-\lau{15-09-2017: TOOD, write more about these. Should be like in the local cap setting.}
+\lau{15-09-2017: TODO, write more about these. Should be like in the local cap setting.}
 
@@
 \item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory seperate from the stack.
-\item[Linking table] A progam may need to call other programs. Other programs are made available in a linking table.
+\item[Linking table] A program may need to call other programs. Other programs are made available in a linking table.
 \end{description}
@@
 The program seals are found on the first address governed by this capability.
-The second address governe by the capability contains a normal read capability that governs the linking table and points to the first address.
+The second address governed by the capability contains a normal read capability that governs the linking table and points to the first address.
 The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
@@
 
-The \emph{peristent data storage} is expected to be accisble through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
-There are two ways for a program to get this data capability: (1) the data part of the code/data part should be the sealed data capability.
+The \emph{peristent data storage} is expected to be accessible through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
+There are two ways for a program to get this data capability: (1) the data part of the code/data pair (for either a callback or a global function in the linking table) should be the sealed data capability.
 (2) During a call, the data capability should be stored on the stack, so it can be retrieved afterwards.
@@
 \]
+\dominique{Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions?
+Is there another function that compiles a single instruction to a list of instructions?}
 
@@
 \label{subsec:capability-opacity}
+
+\dominique{Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
 This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions.
@@
   \end{itemize}
-\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a = a \text{ and } b = b$)
+\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a \rightarrow a \text{ and } b \rightarrow b$)
+  \dominique{I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
   \begin{itemize}
@@
   \item Make the regions more general
-  \item For now, we have not done this: instead we stick with the transition systems we have and have to regions governing the stack reflecting the CSF work.
+  \item For now, we have not done this: instead we stick with the transition systems we have and have two regions governing the stack reflecting the CSF work.
   \end{itemize}

### 2017-09-27 esop-2018/paper.tex
esop folder

--- 2017-09-11 popl-2018/paper.tex
@@
 %% For double-blind review submission
-\documentclass[acmsmall,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true}
+\documentclass{llncs}
 %% For single-blind review submission
@@
 \usepackage{cite}
-\usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{amsmath,amsfonts,amssymb}
 \usepackage{thmtools}
@@
 
-%\def\IEEEbibitemsep{0pt}
-
-% \makeatletter
-% \setlength\mpr@andskip{1em}
-% \def\mpr@lineskip{5em}
-% \def\MathparLineskip{\mpr@lesslineskip}
-% \makeatother
-
-% \newlength{\oldmathindent}
-% \newenvironment{withmathindent}[1]{\setlength{\oldmathindent}{\mathindent}\setlength{\mathindent}{#1}}{\setlength{\mathindent}{\oldmathindent}}
-
-\makeatletter\if@ACM@journal\makeatother
-%% Journal information (used by PACMPL format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{1}
-\acmArticle{1}
-\acmYear{2017}
-\acmMonth{1}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\else\makeatother
-%% Conference information (used by SIGPLAN proceedings format)
-%% Supplied to authors by publisher for camera-ready submission
-\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
-\acmYear{2017}
-\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-\fi
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission
-\setcopyright{none}             %% For review submission
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2017}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
 \begin{document}
@@
 
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{Lau~Skorstengaard}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-  \department{Department of Computer Science}              %% \department is recommended
-  \institution{Aarhus University}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{Aarhus}
-  \state{State1}
-  \postcode{8210}
-  \country{Denmark}
-}
-\email{lau@cs.au.dk}          %% \email is recommended
-
-\author{Dominique~Devriese}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-  \department{iMinds-DistriNet}              %% \department is recommended
-  \institution{KU Leuven}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{-}
-  \state{-}
-  \postcode{-}
-  \country{Belgium}
-}
-\email{dominique.devriese@cs.kuleuven.be}          %% \email is recommended
-
-\author{Lars~Birkedal}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-  \department{Department of Computer Science}              %% \department is recommended
-  \institution{Aarhus University}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{Aarhus}
-  \state{State1}
-  \postcode{8210}
-  \country{Denmark}
-}
-\email{birkedal@cs.au.dk}          %% \email is recommended
-
-%% Paper note
-%% The \thanks command may be used to create a "paper note" ---
-%% similar to a title note or an author note, but not explicitly
-%% associated with a particular element.  It will appear immediately
-%% above the permission/copyright statement.
-%\thanks{with paper note}                %% \thanks is optional
-                                        %% can be repeated if necesary
-                                        %% contents suppressed with 'anonymous'
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
+\bibliographystyle{splncs}
+
+% Authors and affiliation
+\author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
+\institute{Aarhus University\email{\{lau,birkedal\}@cs.au.dk}
+\and iMinds-DistriNet, KU Leuven\email{dominique.devriese@cs.kuleuven.be}}
+
+\maketitle
+
 \begin{abstract}
@@
 
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-% \begin{CCSXML}
-% <ccs2012>
-% <concept>
-% <concept_id>10011007.10011006.10011008</concept_id>
-% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-% <concept_significance>500</concept_significance>
-% </concept>
-% <concept>
-% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-% <concept_significance>300</concept_significance>
-% </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Software and its engineering~General programming languages}
-% \ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords is optional
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
 \section{Introduction}
@@
 machine code rewriting
-(e.g.,~\citet{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
-verification (e.g.,~\citet{morrisett_system_1999}), virtual machines with a
-native stack (e.g.,~the JVM~\citep{lindholm_java_2014}) or
-randomization~\citep{forrest_building_1997}. The latter techniques essentially
+(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
+verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a
+native stack (e.g.,~the JVM~\cite{lindholm_java_2014}) or
+randomization~\cite{forrest_building_1997}. The latter techniques essentially
 emulate protection techniques on existing hardware, at the cost of performance,
@@
 the hardware level. They are based on old ideas
-(e.g.~\citet{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
+(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
 but have recently received renewed interest; in particular, the CHERI
@@
 challenges like backwards compatibility and realistic OS
-support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
+support~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
 machines tag every word (in the register file and in memory) to
@@
 A similar problem is the enforcement of control-flow correctness on capability
-machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
+machines. An interesting approach is taken in CheriBSD~\cite{Watson2015Cheri}:
 the standard contiguous C stack is split into a central, trusted stack, managed
@@
   capability machine (a form of capability
-  safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
+  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
   guarantees offered for local capabilities. It is very general and not tied to
@@
   local capabilities using a variant of Dreyer et. al.'s public and private
-  future worlds~\citep{Dreyer:jfp12}, which they used to express well-bracketedness
+  future worlds~\cite{Dreyer:jfp12}, which they used to express well-bracketedness
   guarantees in a lambda calculus. The logical relation and the
@@
 For reasons of space, some details and all proofs have been omitted; please
-refer to the technical appendix~\citep{technical_appendix} for
+refer to the technical appendix~\cite{technical_appendix} for
 those.\dominique{upload TA as supplementary material instead?}
@@
 like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
-both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and
-CHERI~\citep{Watson2015Cheri}. To avoid uninteresting details, we assume an
+both the M-Machine~\cite{Carter:1994:HSF:195473.195579} and
+CHERI~\cite{Watson2015Cheri}. To avoid uninteresting details, we assume an
 infinite address space and unbounded integers.
@@
     \hline
-    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{w})$&$w = \memreg(r_2)$\\
+    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{w})$&$r_2 \in \Regs \Rightarrow w = \memreg(r_2)$ and $r_2 \in \ints \Rightarrow w = r_2$\\
     \hline
@@
 but it is complicated by the presence of \emph{enter capabilities}, modeled
-after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
+after the M-Machine's~\cite{Carter:1994:HSF:195473.195579}. Enter capabilities
 cannot be used to read, write or execute and their address and range cannot be
@@
 instruction is complicated by the presence of \emph{local capabilities}, modeled
-after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
+after the ones in the CHERI processor~\cite{Watson2015Cheri}. Basically, local
 capabilities are special in that they can only be kept in registers, i.e.\ they
@@
 certain specification. Malloc and linking tables are described further in the
-next section, but we refer to the technical appendix~\citep{technical_appendix}
+next section, but we refer to the technical appendix~\cite{technical_appendix}
 for full details.
@@
 
-The technical appendix~\citep{technical_appendix} contains detailed descriptions
+The technical appendix~\cite{technical_appendix} contains detailed descriptions
 of all the macros.
@@
 types or higher-order ML-style references, and a standard solution is
-to use step-indexing \citep{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
+to use step-indexing \cite{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
 certain number of interaction steps. All words will be considered
@@
 notion of evolvable worlds makes our logical relation into a \emph{Kripke}
-logical relation \citep{pitts_operational_1998}.
+logical relation \cite{pitts_operational_1998}.
 
@@
 high-level
-languages~\citep{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
+languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
 imposed by permanent regions stay in place indefinitely. Any capability, local
@@
 region. Masking a region like this goes back to earlier work of
-\citet{Ahmed2004semantics} and was also used by
-\citet{Thamsborg:2011:KLR:2034773.2034831}.
+\cite{Ahmed2004semantics} and was also used by
+\cite{Thamsborg:2011:KLR:2034773.2034831}.
 
@@
 requirement in Kripke models and the following lemma uses the method of
-\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
+\cite{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
 constructing them. The formulation of the lemma is technical, so we recommend
@@
 them only verbally, but precise definitions can be found in the technical
-appendix~\citep{technical_appendix}.
+appendix~\cite{technical_appendix}.
 
@@
 attempted, like loading from a capability without read permission. This is
-similar to \citet{Devriese:2016ObjCap}'s logical relation for an untyped
+similar to \cite{Devriese:2016ObjCap}'s logical relation for an untyped
 language, but unlike typical logical relations for typed languages, which
@@
 guarantees offered by the capability machine, an instance of a general property
-called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}. To
+called capability safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}. To
 understand this, consider that the theorem says the capability
@@
 macros. The interested reader can find all the technical details in
-the technical appendix~\citep{technical_appendix}.
+the technical appendix~\cite{technical_appendix}.
 
@@
 pointer\footnote{These assumptions are kept intentionally vague for brevity.
-  Full statements are in the technical appendix~\citep{technical_appendix}.}.
+  Full statements are in the technical appendix~\cite{technical_appendix}.}.
 Furthermore
@@
 is a faithful translation of a tricky example known from the literature (known
-as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists
+as the awkward example)~\cite{pitts_operational_1998,Dreyer:jfp12}. It consists
 of two parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
@@
 Formulations in terms of a control-flow graph, like the one by
-\citet{abadi_control-flow_2005}, do not take into account temporal properties,
+\cite{abadi_control-flow_2005}, do not take into account temporal properties,
 like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
@@
 prove that compiling such code to our machine using our calling convention is
-fully abstract~\citep{abadi_protection_1998}.
+fully abstract~\cite{abadi_protection_1998}.
 
@@
 properties are more restrictive and less easy to use for reasoning,
-\citet{Devriese:2016ObjCap} have demonstrated how a logical relation like
+\cite{Devriese:2016ObjCap} have demonstrated how a logical relation like
 ours can be adapted to also support them, by quantifying the logical relation
@@
 existing pattern called biorthogonality. Biorthogonality is a pattern for
-defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998}
+defining logical relations~\cite{krivine_classical_1994,pitts_operational_1998}
 in terms of an observation relation of safe configurations (like we do). The
@@
 
-Interestingly, \citet{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
+Interestingly, \cite{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
 Kripke logical relation for an assembly language (for reasoning about correct
@@
 code that relies on this (particularly, the original awkward example),
-\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
+\cite{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
 contexts using Kripke worlds with public and private future world relations.
@@
 local capabilities are quite similar to a feature proposed for the high-level
-language Scala: \citet{osvald_gentrification_2016}'s second-class or local
+language Scala: \cite{osvald_gentrification_2016}'s second-class or local
 values. They are a kind of values that can be provided to other code for
@@
 
-Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to
-\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
+Capability machines originate with \cite{Dennis:1966:PSM:365230.365252} and we refer to
+\cite{Levy1984capability} and \cite{Watson2015Cheri} for an overview
 of previous work. The capability machine formalized in
@@
 model, modeled mainly after the
-M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers
+M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter pointers
 resemble the M-Machine's) and
-CHERI~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
+CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
 memory and local capabilities resemble CHERI's). The latter is a
@@
 correctness of control flow. This includes particularly the line of work on
-\emph{control-flow integrity}~\citep{abadi_control-flow_2005}. Those use a quite
+\emph{control-flow integrity}~\cite{abadi_control-flow_2005}. Those use a quite
 different attacker model than us: they assume an attacker that is not able to
@@
 secure compilation
-result \citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
+result \cite{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
 differs from theirs in that we use a different form of low-level security
@@
 recursive worlds
-\citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
-closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
+\cite{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
+closely related to the one used by \cite{Devriese:2016ObjCap} to formulate
 capability safety in a high-level JavaScript-like lambda calculus. Our
@@
 from previous work, like Kripke worlds with regions containing state transition
-systems \citep{Ahmed:popl09}, public/private future worlds \citep{Dreyer:jfp12}
+systems \cite{Ahmed:popl09}, public/private future worlds \cite{Dreyer:jfp12}
 (see Section~\ref{sec:discussion} for a discussion), and biorthogonality
-\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
-
-\citet{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
+\cite{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
+
+\cite{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
 of object capability patterns. The logic is based on
-Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
+Iris~\cite{iris,iris2,iris3}, a state of the art higher-order concurrent
 separation logic and is formalized in Coq, building on the Iris Proof Mode for
-Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
+Coq~\cite{ipm}. OCPL gives a more abstract and modular way of proving
 capability safety for a lambda-calculus (with concurrency) compared to the
-earlier work by \citet{Devriese:2016ObjCap}. In the future we would also like to
+earlier work by \cite{Devriese:2016ObjCap}. In the future we would also like to
 investigate a new program logic for reasoning about capability safety for our
@@
 likely that we could get inspiration from earlier work on logics for assembly
-programming languages, such as XCAP~\citep{xcap}.
+programming languages, such as XCAP~\cite{xcap}.
 
@@
 and uses it to prove a compartmentalization
-result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
+result~\cite{akram_el-korashy_formal_2016} (not implying control-flow
 correctness). He also adapts control-flow integrity (see above) to the machine

### 2017-09-27 fully-abstract-comp/technical-report.tex
program definition (work in progress)

--- 2017-09-26 fully-abstract-comp/technical-report.tex
@@
 % Math packages
-\usepackage{amsmath,amsfonts,amssymb}
+\usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{thmtools}
 \usepackage{array}
@@
 \newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+
+%%% Math environments
+\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
+\declaretheorem[numbered=yes,name=Theorem,qed=$\blacksquare$]{theorem}
+\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
+\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}
 
@@
 \newcommand{\stk}{\var{stk}}
+\newcommand{\link}{\var{link}}
 \newcommand{\stkf}{\stk_{\var{frame}}}
@@
 %\newcommand{\rxo}{\plainperm{rxo}}
-%\newcommand{\ro}{\plainperm{ro}}
 \newcommand{\readonly}{\plainperm{r}}
+\newcommand{\ro}{\readonly}
 \newcommand{\noperm}{\plainperm{0}}
@@
 
+\newcommand{\range}[1]{\plainfun{range}{#1}}
+
+\title{Technical Report}
+
 \begin{document}
+\maketitle
+\tableofcontents
 \section{The two capability machines}
@@
 
+\subsubsection{Useful definitions}
+\begin{definition}
+  For a capability $c=((\_,\_),\baddr,\eaddr,\_)$ we say it has range $[\baddr,\eaddr]$ and we define
+  \[
+    \range{c} = [\baddr,\eaddr]
+  \]
+  Similarly for seals and stack pointers:
+  \[
+    \range{\seal{\_,\sigma_\baddr,\sigma_\eaddr,\_}} = [\sigma_\baddr,\sigma_\eaddr]
+  \]
+  and
+  \[
+    \src{\range{\stkptr{\_,\baddr,\eaddr,\_}} = [\baddr,\eaddr]}
+  \]
+\end{definition}
+
 \subsection{Syntax}
@@
 (I think it may have been from when we had call and return instructions?)}
-\dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
 There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
@@
 \subsubsection{Notes}
-Generally:
+Genrally:
 \begin{itemize}
@@
                                           \decPerm{n} \sqsubseteq \perm \tand \\
-                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\Meget
                                           w_2 = \linCons{\Phi(r_2)}
@@
                                   \begin{cases}
+                                    \updPcAddr{}\left(
                                     \arraycolsep=0pt
@@
                                       \Phi&\updReg{r_1}{\vsc}\\
-                                           \updReg{r_2}{w_2}
-                                    \end{array}
+                                          &\updReg{r_2}{w_2}
+                                    \end{array}\right)
 &
@@
 \noindent\textbf{split and splice}\\
-We would like splice and split to have the following properties
+We would like splice and split to have following properties
 \begin{enumerate}
@@
 
-\subsection{Compiler}
+\subsection{Programs and contexts}
+\begin{definition}
+  \label{def:self-contained}
+  \begin{itemize}
+  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$
+    are self-contained iff
+  \[
+    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom{\ms}
+  \]
+  and
+  \begin{multline*}
+    \forall a \in \dom(\ms) \ldotp\\
+    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom{\ms}
+  \end{multline*}
+  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$ are self contained with slack $\ms_s$ iff
+  \[
+    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom{\ms \cup \ms_s}
+  \]
+  and
+  \begin{multline*}
+    \forall a \in \dom(\ms) \ldotp\\
+    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom{\ms \cup \ms_s}
+  \end{multline*}
+\end{itemize}
+\end{definition}
+
+\subsubsection{Programs}
+Here we define what consitutes a program.
+
+\begin{definition}
+  \label{def:program}
+  \lau{27-09-2017: The ``somehow''s need to be fixed.}
+  A program consists of two sealed capabilities, $\vsc_\code = \sealed{c_\code,\sigma}$ and $\vsc_\data = \sealed{c_\data,\sigma}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
+  \begin{itemize}
+  \item $c_\code = ((\rx,\lin_\code),\baddr_\code,\eaddr_\code,\aaddr_\code)$ where $\aaddr = \baddr + 3$ and $[\baddr_\code,\eaddr_\code]\subseteq \dom(\ms)$
+  \item $c_\data = ((\rw,\lin_\data),\baddr_\data,\eaddr_\data,\baddr_\data)$ where $[\baddr_\data,\eaddr_\data] \subseteq \dom(\ms)$ and $[\baddr_\code,\eaddr_\code] \cap [\baddr_\data,\eaddr_\data] = \emptyset$
+  \item $\ms$ and $[c_\code,c_\data]$ are self-contained with $\ms_\link$ as slack.
+  \item $\ms(\baddr_\code) = \seal(\normal,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr)$ where these are somehow the sufficient amount of seals.
+  \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
+  \item $\ms(\baddr_\code+2) = \dots$ where $\dots$ is somehow the stack height.
+  \end{itemize}
+\end{definition}
+\lau{27-09-2017: At the moment, we do not have the proper support to let the code capability be linear (in the sense that you cannot write a sensible program) - maybe we should just require the code capability to be normal.}
+
+\subsubsection{Contexts}
+
+\section{Compiler}
 \[

### 2017-09-27 esop-2018/paper.tex
fixed (most) things, so they fit with margin

@@
 \newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
-\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
+\newcommand{\execCond}[1]{\plainfun{execCondition}{#1}}
 \newcommand{\entryCond}[1]{\plainfun{enterCondition}{#1}}
@@
 \[\arraycolsep=1.4pt
-  \begin{array}{rlrcl p{5cm}rlrcl}
-    \addr   &\in& \Addrs &\isdef& \nats &\phantom{makesomero}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
-    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words\\
+  \begin{array}{rlrcl l rlrcl}
+    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
+    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
     \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
@@
                  \store{\hv}{\rv} \mid \plus{\lv}{\rv}{\rv} \mid 
-                 \minus{\lv}{\rv}{\rv} \mid 
-                 \lt{\lv}{\rv}{\rv} \mid \\
-          &   &  \lea{\lv}{\rv} \mid 
+                 \minus{\lv}{\rv}{\rv} \mid \\
+          &   &  \lt{\lv}{\rv}{\rv} \mid
+                \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid
@@
                  \isptr{\lv}{\rv} \mid 
-                 \getp{\lv}{\lv} \mid 
-                 \getl{\lv}{\lv} \mid 
-                 \getb{\lv}{\lv} \mid \\
-          &   &  \gete{\lv}{\lv} \mid 
+                 \getl{\lv}{\lv} \mid \\
+          &   &  \getp{\lv}{\lv} \mid 
+                 \getb{\lv}{\lv} \mid
+                 \gete{\lv}{\lv} \mid 
                  \geta{\lv}{\lv} \mid 
@@
         &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.2cm]
-        &\quad \forall r \in \activeReg{W} \ldotp \exists H,s \ldotp W(r) = (\_,s,\_,\_,H) \text{ and } \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
+        &\quad \forall r \in \activeReg{W} \ldotp\\
+        &\quad \quad \exists H,s \ldotp W(r) = (\_,s,\_,\_,H) \text{ and } \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
       \end{aligned}\right.
@@
           & \;\; %\Rightarrow
-            \exists W' \futurestr W, \ms_r, \ms' \ldotp \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \heapSat[\hs']{n-i}{W'}
+            \exists W' \futurestr W, \ms_r, \ms' \ldotp \\
+          & \;\;\;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \heapSat[\hs']{n-i}{W'}
         \end{aligned}\right\}\\
@@
         \begin{multlined}
-          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp 
+          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
           \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W)
         \end{multlined} \right\} \\
-      & \\
+      & \\[-0.2cm]
       &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
@@
              \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
-             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in
-            \execCond{}(\gl)(W)
+             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \\
+             \hfill \in \execCond{}(\gl)(W)
             \end{array}
@@
   & \readCond{}(\gl)(W) =  \left\{ \npair{(\start,\addrend)} \; \middle| \;
-       \exists r \in \var{localityReg}(g,W) \ldotp 
-       \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
+    \begin{array}{l}
+       \exists r \in \var{localityReg}(g,W) \ldotp \\
+\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
+    \end{array}
     \right\}\\
@@
     \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp W(r) \text{ is address-stratified and} \\
+      & \exists r \in \var{localityReg}(g,W) \ldotp \\
+      & W(r) \text{ is address-stratified and} \\[-0.2cm]
       & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
@@
     \left\{
-      \npairP{(\var{perms}, \start,\addrend)}
+      \npairP{(\var{P}, \start,\addrend)}
      \; \middle| \;
     \begin{aligned}
-      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
+      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{P} \ldotp \\
       & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
@@
     \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
-    \forall n' < n \ldotp \forall W' \future W \ldotp \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
+    \begin{array}{l}
+      \forall n' < n \ldotp \forall W' \future W \ldotp\\
+      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
+    \end{array}
     \right\} \\

### 2017-09-28 fully-abstract-comp/technical-report.tex
spelling

@@
 \lau{14-09-2017: Insert the relation.
-My impression is that the the words specific to the source machine should have the same "type" as their implementaion at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation).
+My impression is that the the words specific to the source machine should have the same "type" as their implementation at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation).
 }
@@
                                           \decPerm{n} \sqsubseteq \perm \tand \\
-                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\Meget
+                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\
                                           w_2 = \linCons{\Phi(r_2)}
@@
 \item No authority amplification - splitting or splicing capabilities should give you no more authority than you already had.
-\item Split should be dual to splice in the sence that a split on a capability followed by a splice of the two resulting capabilities should yield the same capability.
+\item Split should be dual to splice in the sense that a split on a capability followed by a splice of the two resulting capabilities should yield the same capability.
 \item Take the addresses governed by a linear capability to be a multiset.
@@
 \lau{19-09-2017: Because $\tsplice{}{}{}$ does not allow empty stacks, it is not ``left inverse'' to $\tsplit{}{}{}{}$ (because of the empty case).
-Intuitively, it is weird that a $\tsplit{}{}{}{}$ followed by a $\tsplice{}{}{}$ does not yeild the same capability.}
+Intuitively, it is weird that a $\tsplit{}{}{}{}$ followed by a $\tsplice{}{}{}$ does not yield the same capability.}
 \begin{align*}
@@
 \item[Stack] A program stores the value of local variables on the stack. 
-\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory seperate from the stack.
+\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory separate from the stack.
 \item[Linking table] A program may need to call other programs. Other programs are made available in a linking table.
@@
 
-The \emph{peristent data storage} is expected to be accessible through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
+The \emph{persistent data storage} is expected to be accessible through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
 There are two ways for a program to get this data capability: (1) the data part of the code/data pair (for either a callback or a global function in the linking table) should be the sealed data capability.
@@
 \subsubsection{Programs}
-Here we define what consitutes a program.
+Here we define what constitutes a program.
 
@@
   \begin{itemize}
-  \item Allows backtranslation to be embedding into source language.
-    \begin{itemize}
-    \item Interpreter for backtranslation is not able to accurately replicate code length.
+  \item Allows back translation to be embedding into source language.
+    \begin{itemize}
+    \item Interpreter for back translation is not able to accurately replicate code length.
     \end{itemize}
@@
 \item Components are memory segments and symbols.
-\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, se example in subsection~\ref{subsec:capability-opacity}).
+\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, see example in subsection~\ref{subsec:capability-opacity}).
   \begin{itemize}
@@
 
-p1 and p2 each accept an (unprotected) closure from the context whose lenght is restricted.
+p1 and p2 each accept an (unprotected) closure from the context whose length is restricted.
 Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
@@
 However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{}{} compiles to.
-We will discuss later that this means that if we backtranslate an adversary piece of code (in the full abstraction proof), then it will be backtranslated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
+We will discuss later that this means that if we back translate an adversary piece of code (in the full abstraction proof), then it will be back translated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
 In other words, the back-translated code never uses call to protect its own stack frame from the code that it passes control to.
@@
 
-Note that in the new setup with linear capabilities instead of local/global ones, there seems to be less opportunity for the trusted code to do stupid things, so it may be that we no lnoger need these sanity conditions, but this is not clear yet.
+Note that in the new setup with linear capabilities instead of local/global ones, there seems to be less opportunity for the trusted code to do stupid things, so it may be that we no longer need these sanity conditions, but this is not clear yet.
 
@@
 In order for this to work, the memory must initially be partitioned into two parts.
-One controlled by a non-linear capability and one controled by a linear capability.
+One controlled by a non-linear capability and one controlled by a linear capability.
 The semantics of the machine would now have to make sure that the linear capability is never aliased.
@@
 This guarantee is maintained by the semantics.
-Linear capabilties cannot be converted to another kind of capability (in other words, this is option one but where linear capabilities cannot be made non-linear).
+Linear capabilities cannot be converted to another kind of capability (in other words, this is option one but where linear capabilities cannot be made non-linear).
 
@@
 \end{enumerate}
-Notice that the two first schemes are incompatible as they would undermine eachother.
+Notice that the two first schemes are incompatible as they would undermine each other.
 
@@
 Further something like write-linear would have the same issue as write-local.
-We would rely on the fact that the write-linear permission limits the space where we might need to look for linear capabilties.
+We would rely on the fact that the write-linear permission limits the space where we might need to look for linear capabilities.
 In order to remove the desired linear capability, we would still need to search through all the places it could be which would be infeasible (like the clear memory segment instruction).}

### 2017-09-28 esop-2018/paper.tex
Improve alignment in a figure.
Fix affiliation.
Save some space by using wrapfig and playing with figure placement.
Save some more space.

@@
 \usepackage{cite}
+\usepackage{wrapfig}
 \usepackage{amsmath,amsfonts,amssymb}
@@
 \institute{Aarhus University\email{\{lau,birkedal\}@cs.au.dk}
-\and iMinds-DistriNet, KU Leuven\email{dominique.devriese@cs.kuleuven.be}}
+\and imec-DistriNet, KU Leuven\email{dominique.devriese@cs.kuleuven.be}}
 
@@
 
-\begin{figure}
+\begin{figure}[p]
 \[\arraycolsep=1.4pt
@@
   \label{fig:syntax}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (7) {$\rwlx$};
-    \node[main node] (8) [below left of=7] {$\readwritel$};
-    \node[main node] (1) [below right of=7] {$\rwx$};
-    \node[main node] (2) [below right of=1] {$\exec$};
-    \node[main node] (3) [below right of=2] {$\entry$};
-    \node[main node] (4) [below left of=1] {$\readwrite$};
-    \node[main node] (5) [below right of=4] {$\readonly$};
-    \node[main node] (6) [below right of=5] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (7) edge (8)
-    (7) edge (1)
-    (8) edge (4)
-    (1) edge (2)
-    (2) edge (3)
-    (2) edge (5)
-    (3) edge (6)
-    (1) edge (4)
-    (4) edge (5)
-    (5) edge (6);
-  \end{tikzpicture}
-
-  \caption{Permission hierarchy}
-  \label{fig:perm-hier}
 \end{figure}
@@
 \end{figure*}
+
+\begin{wrapfigure}{R}{.3\textwidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (7) {$\rwlx$};
+    \node[main node] (8) [below left of=7] {$\readwritel$};
+    \node[main node] (1) [below right of=7] {$\rwx$};
+    \node[main node] (2) [below right of=1] {$\exec$};
+    \node[main node] (3) [below right of=2] {$\entry$};
+    \node[main node] (4) [below left of=1] {$\readwrite$};
+    \node[main node] (5) [below right of=4] {$\readonly$};
+    \node[main node] (6) [below right of=5] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (7) edge (8)
+    (7) edge (1)
+    (8) edge (4)
+    (1) edge (2)
+    (2) edge (3)
+    (2) edge (5)
+    (3) edge (6)
+    (1) edge (4)
+    (4) edge (5)
+    (5) edge (6);
+  \end{tikzpicture}
+
+  \caption{Permission hierarchy}
+  \label{fig:perm-hier}
+\end{wrapfigure}
 
@@
 \subsection{Worlds}
-\begin{figure}
+\begin{figure}[tbp]
   \centering
-  \includegraphics{w11}
+  \resizebox{6cm}{!}{\includegraphics{w11}}
+
   \caption{An illustration of local/global capabilities and temporary/permanent
@@
   \begin{align*}
-  & \readCond{}(\gl)(W) =  \left\{ \npair{(\start,\addrend)} \; \middle| \;
+   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
     \begin{array}{l}
@@
     \right\}\\
-  & \writeCond{}(\iota,\gl)(W) =  \left\{
+   \writeCond{}(\iota,\gl)(W) &=  \left\{
     \npair{(\start,\addrend)}
@@
     \end{aligned} \right\}\\
-  & \execCond{}(\gl)(W) = 
+   \execCond{}(\gl)(W) &= 
     \left\{
@@
     \end{aligned} \right\} \\
-  & \entryCond{}(\gl)(W) = 
+   \entryCond{}(\gl)(W) &= 
     \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
@@
   \centering
-  \begin{minipage}[t]{5.05cm}
+  \begin{minipage}[t]{.37\linewidth}
   \begin{lstlisting}
@@
      prepstk $r_\stk$
+     ${\tiny\itshape\text{(continues in next column)}}$
+\end{lstlisting}
+  \end{minipage}%
+  \begin{minipage}[t]{.37\linewidth}
+\begin{lstlisting}
+     ${\tiny\itshape\text{(continued from previous column)}}$
      store $x$ 0
@@
 \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}[t]{3.4cm}
+  \end{minipage}%
+  \begin{minipage}[t]{.25\linewidth}
   \begin{lstlisting}

### 2017-09-28 fully-abstract-comp/technical-report.tex
Added responses to Dominique's comments

@@
 \dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
+\lau{28-09-2017: Why would you like this?
+If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail.
+We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
 \dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
+\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part.
+Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}.
+This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame.
+Makes sense?}
 The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
@@
     n & \in & \nats \\
-    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
+    \src{r} &\in &  \RegName\\
     \src{\rn} &\defbnf & \src{r} \mid n \\
@@
 \dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
+\lau{28-09-2017: I can see from the commit history that this was something when we had push/pop instructions as well as a return instruction. I have removed the restriction on $\rstk$.}
 There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
@@
 
-\dominique{I think we need to do something about tail calls in the source language?}
+\dominique{27-09-2017: I think we need to do something about tail calls in the source language?}
+\lau{27-09-2017: In what sense? Do you want to allow them (I do not thinkn they are allowed at the moment?)}
 
@@
 \]
-\dominique{Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions?
+\dominique{27-09-2017: Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions?
 Is there another function that compiles a single instruction to a list of instructions?}
+\lau{28-09-2017: Yes, that was the thought at the moment, but this may very well change soon.
+This compiler would then use a compiler that compiles single instructions.
+That said, this part was added a long time ago and for some reason (some details about the compilation probably did not make sense to me at the moment), I did not complete it.
+I think the compiler will probably change to be from programs to programs, and a programs is not just going to be a list of instructions.}
 
@@
 
-\dominique{Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
+\dominique{27-09-2017: Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
+\lau{28-09-2017: It seems to me that this is indeed solved by the variable length instructions.
+That said, I would like to keep it around (maybe change it so it becomes an argument for variable length instructions).}
 This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions.
@@
 \item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a \rightarrow a \text{ and } b \rightarrow b$)
-  \dominique{I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
+  \dominique{27-09-2017: I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
+  \lau{28-09-2017: I think the next three points is the answer.
+We basically wanted a formalism in which we could design one region that would govern the stack.
+I do not remeber why it was made possible with PORs.}
   \begin{itemize}

### 2017-09-28 fully-abstract-comp/technical-report.tex
adjusted splice semantics slightly

@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = ((\perm,\lin),\baddr_2,\eaddr_2,\aaddr_2) \tand \\
-                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\_) \tand \\
+                                  \Phi(r_2) = ((\perm,\lin),\baddr_2,\eaddr_2,\_) \tand \\
+                                  \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\aaddr_3) \tand \\
                                   \perm \neq \enter \tand \\
                                   \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
-                                  c = ((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_2) \tand\\
+                                  c = ((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3) \tand\\
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2},\sigma} \tand \\
-                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3},\_} \tand \\
+                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2},\_} \tand \\
+                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3},\sigma} \tand \\
                                   \sigma_{\eaddr,2}+1 = \sigma_{\baddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand\\
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\aaddr_2} \tand \\
+                                  \Phi(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_} \tand \\
                                   \Phi(r_3) = \stkptr{\perm,\baddr_3,\eaddr_3,\aaddr_3} \tand \\
@@
                                   \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
-                                  c = \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_2} \\
+                                  c = \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_3} \\
                                 \end{array}\\
@@
                                                    \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
-                                                   \baddr_\stk \leq \aaddr_\stk < \eaddr_\stk \tand \\
-                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\baddr_\stk,\aaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
-                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\aaddr_\stk + 1,\eaddr_\stk]} \tand \\
-                                                   c_\stk = \stkptr{\aaddr_\stk+1,\eaddr_\stk,\aaddr_\stk+1} \tand \\
+                                                   \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
+                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
+                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr,\aaddr_\stk - 1]} \tand \\
+                                                   c_\stk = \stkptr{\baddr,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
                                                    \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
@@
 
-One way to solve this is to make it publicly known what the end address of the stack is and always check that we receive a part of the stack that goes all the way to the end.
-In the above example (using the stack parts in parenthesis) this would cause the sequence of calls to fail the second time we are called, because we do not receive a stack that goes all the way to the top of the stack.
+One way to solve this problem is to make the "top addess" of the stack known. There are many ways to do this, but we have chosen to do the following:
+\begin{itemize}
+\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
+\item The base address of the stack capability is a fixed address, so in the semantice, it will be expressed as a constant that is publicly known.
+\end{itemize}
 

### 2017-09-28 fully-abstract-comp/technical-report.tex
changes to the semantics related to our wednesday discussions.

@@
 \newcommand{\aendbnf}{\shareddom{end}}
+\newcommand{\rbnf}{\shareddom{r}}
 %\newcommand{\glbnf}{\shareddom{gl}}
@@
 \newcommand{\calllen}{\constant{call\_len}}
+\newcommand{\stkb}{\constant{stk\_base}}
 
@@
     &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
-    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
+    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
     &\Reg & \defeq & \RegName \fun \Word\\
@@
 \item $\cbnf$ defines domain $\Caps$
+\item $\rbnf$ defines the finite set $\RegName$.
 \item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
@@
 The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) caller's private stack, and 3) the address of the old stack pointer.
+
+In both languages, the base address of the stack is known (the stack grows downwards in memory, so the base address marks the end of the stack).
+In the target language, this will be the base address of some linear capability.
+The base address can be any address on the machine, but it will have to remain the same.
+We write this constant as
+\[
+  \stkb
+\]
 
@@
                                          \Phi(r_2) = \retptrd(\sigma) \tand \\
-                                         \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_} \tand \\
+                                         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
                                          \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                          \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
-                                         \dom(\ms_{\stk,\priv}) = [\baddr_{\stk,\priv},\baddr_\stk - 1] \tand\\
-                                         \ms_{\stk,\priv}(\baddr_\stk-1) = \eaddr_\stk \tand \\
-                                         c_\stk = \stkptr{\rw,\baddr_{\stk,\priv},\eaddr_\stk,\aaddr_\stk} \tand\\
+                                         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
+                                         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
@@
                                                    \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
-                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr,\aaddr_\stk - 1]} \tand \\
-                                                   c_\stk = \stkptr{\baddr,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
+                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk - 1]} \tand \\
+                                                   c_\stk = \stkptr{\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
                                                    \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\

### 2017-09-28 esop-2018/paper.tex
reverted the calling convention description to its previous unreadable state.

@@
 \textbf{Calling convention} With the security measures introduced and motivated,
-let us summarize our proposed calling convention:
-\begin{description}[font=\normalfont\itshape]
-\item[At program start-up] A local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local capabilities.
-\item[Before returning to the adversary] Clear non-return-value registers. Clear
-  the part of the stack we had access to (not just the part we used).
-\item[Before invoking the adversary] Push activation record to the stack. Create
-  return pointer as local $\entry$-capability to the instructions in the record.
-  Restrict the stack capability to the unused part and clear it. Clear
-  non-argument registers.
-\item[Before invoking an adversary callback] Make sure callback is global.
-\item[When invoked by an adversary] Make sure received stack pointer has permission $\rwlx$.
-\end{description}
+let us summarize our proposed calling convention: \textit{At program start-up} A
+local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local
+capabilities. \textit{Before returning to the adversary} Clear non-return-value
+registers. Clear the part of the stack we had access to (not just the part we
+used). \textit{Before invoking the adversary} Push activation record to the
+stack. Create return pointer as local $\entry$-capability to the instructions in
+the record.  Restrict the stack capability to the unused part and clear
+it. Clear non-argument registers. \textit{Before invoking an adversary callback}
+Make sure callback is global. \textit{When invoked by an adversary} Make sure
+received stack pointer has permission $\rwlx$.
 
@@
      prepstk $r_\stk$
-     ${\tiny\itshape\text{(continues in next column)}}$
+     ${\text{\tiny\itshape\text{(continues in next column)}}}$
 \end{lstlisting}
@@
 \begin{lstlisting}
-     ${\tiny\itshape\text{(continued from previous column)}}$
+     ${\text{\tiny\itshape\text{(continued from previous column)}}}$
      store $x$ 0

### 2017-09-28 esop-2018/paper.tex
added more about the stack end point check
Changed a value in esop paper.
layout changes.

@@
 \subsection{Worlds}
-\begin{figure}[tbp]
+\begin{wrapfigure}{R}{0.5\textwidth}
   \centering
@@
 
-  \caption{An illustration of local/global capabilities and temporary/permanent
-    regions. The table represents a memory and the colored fields are regions
-    governing a memory range. Global capabilities cannot depend on temporary
+  \caption{The relation between local/global capabilities and temporary/permanent
+    regions. The colored fields are regions governing parts of memory. Global capabilities cannot depend on temporary
     regions.}
   \label{fig:cap-world}
-\end{figure}
+\end{wrapfigure}
 % \dominique{Perhaps we could first try to convey the intuition of a permanent
@@
     \end{align*}
-  and for $W, W' \in \Wor$
-  \begin{align*}
+  and for $W, W' \in \Wor\ldotp \quad  
+  \begin{aligned}
     W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
     W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
-  \end{align*}
+  \end{aligned}$
 \end{theorem}
@@
 The \emph{private future world} relation $W' \futurestr W$ is defined similarly,
-using a private future region relation.
-
-The \emph{public future} region relation is the simplest. It satisfies
+using a private future region relation. The \emph{public future} region relation is the simplest. It satisfies
 % \dominique{is defined by?} 
@@
 public future world relations allows us to reinstate a region that has been revoked
-earlier.
-
-The \emph{private future region} relation satisfies:
+earlier. The \emph{private future region} relation satisfies:
 \begin{mathpar}
@@
     \text{ iff }\quad \left\{\begin{aligned}
-        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.2cm]
+        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.35cm]
         &\quad \forall r \in \activeReg{W} \ldotp\\
@@
   and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
-  Then if
-  \[
-  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
-  \]
+  Then if $
+  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
   then $m'(\flag) = 0$.

### 2017-09-28 esop-2018/paper.tex
removed 'redundant' definitions in the value relation that did not show new permission based conditions.
more small gains

@@
 
+\newcommand{\sectionname}{Sec.}
 \newcommand{\forcenewline}{$\phantom{v}$\\}
@@
 
-\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
-\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
-\newcommand{\execCond}[1]{\plainfun{execCondition}{#1}}
-\newcommand{\entryCond}[1]{\plainfun{enterCondition}{#1}}
+\newcommand{\readCond}[1]{\plainfun{readCond}{#1}}
+\newcommand{\writeCond}[1]{\plainfun{writeCond}{#1}}
+\newcommand{\execCond}[1]{\plainfun{execCond}{#1}}
+\newcommand{\entryCond}[1]{\plainfun{enterCond}{#1}}
 
@@
   capabilities, and its operational semantics
-  (Section~\ref{sec:capab-mach-with}).
+  (\sectionname~\ref{sec:capab-mach-with}).
 \item We define a novel calling convention enforcing control-flow correctness and
-  encapsulation of stack frames (Section~\ref{sec:stack-and-return-pointer}). It
+  encapsulation of stack frames (\sectionname~\ref{sec:stack-and-return-pointer}). It
   relies solely on local capabilities and does not require OS support (like a
@@
   fundamental theorem expressing capability safety
-  are presented in Section~\ref{sec:logical-relation}.
+  are presented in \sectionname~\ref{sec:logical-relation}.
 \item We demonstrate our results by applying them to challenging examples,
@@
   control-flow correctness guarantees in the presence of cross-component
-  function pointers (Section~\ref{sec:examples}). The examples demonstrate both
+  function pointers (\sectionname~\ref{sec:examples}). The examples demonstrate both
   the power of our formulation of capability safety and our calling convention.
@@
 requirement is the most problematic aspect of our calling convention for
-performance, but see Section~\ref{sec:discussion} for how this might be
+performance, but see \sectionname~\ref{sec:discussion} for how this might be
 mitigated.
@@
 % \lau{If we need space, then maybe one of the examples is sufficient.}
-% \lau{Both of these are good to have their as the two last examples of Section~\ref{sec:examples} use them.}
+% \lau{Both of these are good to have their as the two last examples of \sectionname~\ref{sec:examples} use them.}
 %
@@
 worlds that has been previously used in the literature (see
-Section~\ref{sec:discussion}). However, temporary regions are new in our setting
+\sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
 and there is an interesting interplay with the recursiveness of the worlds: for
@@
 % world-dependent. 
-As described in Section~\ref{sec:formalizing-guarantees}, regions define sets of
+As described in \sectionname~\ref{sec:formalizing-guarantees}, regions define sets of
 safe memory segments, but this set may itself be world-dependent. In other
@@
         \union \\
-        & \left\{
-            \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-          \right\}
-        \union \\
+        % & \left\{
+        %     \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+        %   \right\}
+        % \union \\
         & \left\{
@@
           \right\} \union \\
-        & \left\{
-            \npair{\stdcap[(\readwritel,\gl)] } \;\middle| \;
-            \begin{array}{l}
-             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
-            \end{array}
-           \right\}
-        \union \\
-        & \left\{
-          \npair{\stdcap[(\exec,\gl)]} \; \middle| \;
-          \begin{array}{l}
-             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-          \end{array}
-           \right\}
-        \union \\
+        % & \left\{
+        %     \npair{\stdcap[(\readwritel,\gl)] } \;\middle| \;
+        %     \begin{array}{l}
+        %      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+        %      \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
+        %     \end{array}
+        %    \right\}
+        % \union \\
+        % & \left\{
+        %   \npair{\stdcap[(\exec,\gl)]} \; \middle| \;
+        %   \begin{array}{l}
+        %      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+        %      \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+        %   \end{array}
+        %    \right\}
+        % \union \\
         & \left\{
@@
         \union \\ 
-        & \left\{
-          \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
-          \begin{array}{l}
-            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-            \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
-            \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-          \end{array}
-          \right\}
-        \union \\
+        % & \left\{
+        %   \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
+        %   \begin{array}{l}
+        %     \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+        %     \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
+        %     \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+        %   \end{array}
+        %   \right\}
+        % \union \\
         & \left\{
@@
              \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
-             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \\
-             \hfill \in \execCond{}(\gl)(W)
+             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
             \end{array}
-          \right\}
+          \right\} \\
+          \union
+        & \dots \small{\text{\itshape and so on for permissions $\readonly$, $\readwritel$, $\exec$, and $\rwx$.}}
       \end{aligned}
@@
     \right\} \\
-  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
-  & \quad\;\; \text{and } \gl = \glob \Rightarrow \future = \futurestr
+   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
   \end{align*}
@@
 configurations are safe according to the intuitions described in
-Section~\ref{sec:formalizing-guarantees}. The definition is found in
+\sectionname~\ref{sec:formalizing-guarantees}. The definition is found in
 \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
@@
 global. Note finally that the step-indexing described in
-Section~\ref{sec:formalizing-guarantees} is effectively at play here, but hidden
+\sectionname~\ref{sec:formalizing-guarantees} is effectively at play here, but hidden
 in the definition of the standard region $\iota^\pwl_{\start,\addrend}$, which
@@
 that they prohibit global write-local capabilities which, as discussed in
-Section~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
+\sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
 local capabilities.
@@
 of the values it has access to. We demonstrate this in
-Section~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
+\sectionname~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
 capabilities that point to arbitrary instructions. The relation between
 Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow
-correctness, will also be shown by example in Section~\ref{sec:examples} as the
+correctness, will also be shown by example in \sectionname~\ref{sec:examples} as the
 examples depend on these properties for correctness.
@@
 correctness properties for challenging program examples. The security
-measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
+measures of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to
 ensure these properties. Since we are dealing with assembly language,
@@
 top of its stack and the assertion would fail. However, the security measures in
-Section~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
+\sectionname~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
 the return pointer is local, so it can only be stored on the stack, but the part
@@
 adversary and an adversary callback invoked by us. As explained in
-Section~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
+\sectionname~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
 that the stack pointer that the closure receives from the adversary has
@@
 
-As explained in Section~\ref{sec:stack-and-return-pointer}, the
+As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
 macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
@@
 security measures of the calling convention described in
-Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
+\sectionname~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
 calling convention without security guarantees. However, most of our security
@@
 \emph{Semantic, but not syntactic properties} The logical relation defined in
-Section~\ref{sec:logical-relation} allows us to reason about capability machine
+\sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
 programs. A limitation w.r.t. previous work is that the logical relation is
@@
 Finally, we summarize how our work relates to previous work. We do not
-repeat the work we discussed in Section~\ref{sec:discussion}.
+repeat the work we discussed in \sectionname~\ref{sec:discussion}.
 
@@
 of previous work. The capability machine formalized in
-Section~\ref{sec:capab-mach-with} is a simple but representative
+\sectionname~\ref{sec:capab-mach-with} is a simple but representative
 model, modeled mainly after the
@@
 will follow a legal path in the control-flow graph. As discussed in
-Section~\ref{sec:discussion}, such a property ignores temporal properties and
+\sectionname~\ref{sec:discussion}, such a property ignores temporal properties and
 seems hard to use for reasoning.
@@
 systems \cite{Ahmed:popl09}, public/private future worlds \cite{Dreyer:jfp12}
-(see Section~\ref{sec:discussion} for a discussion), and biorthogonality
+(see \sectionname~\ref{sec:discussion} for a discussion), and biorthogonality
 \cite{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

### 2017-09-28 esop-2018/paper.tex
removed part of the example section, namely the world construction for the last proof.

@@
 \texttt{\footnotesize{f4}} returns.
-
+By utilizing the reinstation mechanism of the future world relation as well as our knowledge of the future worlds used, we can construct a world which the adversary's invariants are preserved.
+The details of this and the proofs of the other lemmas are found in the technical appendix\cite{technical_appendix}.
+
+\begin{comment}
 In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
@@
 \end{figure}
+\end{comment}
 \begin{comment}

### 2017-09-29 esop-2018/paper.tex
removed syntactic/semantic discussion from esop paper.

@@
 \\\\
-\noindent \textbf{Reasoning about capability machine programs}\\
-\emph{Semantic, but not syntactic properties} The logical relation defined in
-\sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
-programs. A limitation w.r.t. previous work is that the logical relation is
-tailored exclusively towards semantic properties, not syntactic ones.
-
-Imagine, for example, that we invoke a block of adversary code in such a way
-that it only ever receives capabilities within a specific range of memory. After
-the code returns, we may try to prove that any capabilities passed back to us in
-the registers are still confined to that range of memory. The property of
-falling in a certain address range is syntactic in the sense that it talks about
-the specific implementation of a higher-order value rather than its behavior,
-like the invariants that are required/preserved when we use it.
-
-Such syntactic properties are hard to prove in our system. For the example
-cited, it would be easy to conclude that the returned values are in the value
-relation (see Figure~\ref{fig:logrel}). This gives us a lot of semantic
-information, like conditions under which they are safe to use and invariants
-that will be preserved when we do, but it does not tell us much about the
-address of the capability. As a very concrete example, capabilities with
-permission $\noperm$ are always in the value relation, irrespective of their
-address. Semantically, this makes perfect sense, since they are always safe
-since they cannot be used for anything anyway. However, it also means we do not
-get syntactic information about them.
-
-For our purposes, this restriction is unproblematic, since we are only
-interested in proving semantic properties (e.g., an assertion will never fail).
-However, in other situations, we may be interested in proving more syntactic
-properties like the ones that are often considered in object capability
-literature: confinement, no authority amplification etc. Although such
-properties are more restrictive and less easy to use for reasoning,
-\cite{Devriese:2016ObjCap} have demonstrated how a logical relation like
-ours can be adapted to also support them, by quantifying the logical relation
-over a custom interpretation of effectful computations and the type of
-references. We expect their solution can be readily adapted to our setting,
-modulo some details (like the fact that we do not just have read-write
-capabilities, but also others).
-% \emph{No program logic} TODO: discuss this? 
-\\\\
+% \noindent \textbf{Reasoning about capability machine programs}\\
+% \emph{Semantic, but not syntactic properties} The logical relation defined in
+% \sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
+% programs. A limitation w.r.t. previous work is that the logical relation is
+% tailored exclusively towards semantic properties, not syntactic ones.
+
+% Imagine, for example, that we invoke a block of adversary code in such a way
+% that it only ever receives capabilities within a specific range of memory. After
+% the code returns, we may try to prove that any capabilities passed back to us in
+% the registers are still confined to that range of memory. The property of
+% falling in a certain address range is syntactic in the sense that it talks about
+% the specific implementation of a higher-order value rather than its behavior,
+% like the invariants that are required/preserved when we use it.
+
+% Such syntactic properties are hard to prove in our system. For the example
+% cited, it would be easy to conclude that the returned values are in the value
+% relation (see Figure~\ref{fig:logrel}). This gives us a lot of semantic
+% information, like conditions under which they are safe to use and invariants
+% that will be preserved when we do, but it does not tell us much about the
+% address of the capability. As a very concrete example, capabilities with
+% permission $\noperm$ are always in the value relation, irrespective of their
+% address. Semantically, this makes perfect sense, since they are always safe
+% since they cannot be used for anything anyway. However, it also means we do not
+% get syntactic information about them.
+
+% For our purposes, this restriction is unproblematic, since we are only
+% interested in proving semantic properties (e.g., an assertion will never fail).
+% However, in other situations, we may be interested in proving more syntactic
+% properties like the ones that are often considered in object capability
+% literature: confinement, no authority amplification etc. Although such
+% properties are more restrictive and less easy to use for reasoning,
+% \cite{Devriese:2016ObjCap} have demonstrated how a logical relation like
+% ours can be adapted to also support them, by quantifying the logical relation
+% over a custom interpretation of effectful computations and the type of
+% references. We expect their solution can be readily adapted to our setting,
+% modulo some details (like the fact that we do not just have read-write
+% capabilities, but also others).
+% % \emph{No program logic} TODO: discuss this? 
+% \\\\
 \noindent\textbf{Logical relation}\\

### 2017-09-29 esop-2018/paper.tex
typo

@@
 such scenarios to guarantee safety and preserve our invariants, but such checks
-may be costly and it is not clear to us whether there are practical scenario's
+may be costly and it is not clear to us whether there are practical scenarios
 where this might be realistic.

### 2017-09-29 esop-2018/paper.tex
Fix a latex compilation issue I seem to have introduced.

--- 2017-09-28 esop-2018/paper.tex
@@
 
-\newcommand{\sectionname}{Sec.}
 \newcommand{\forcenewline}{$\phantom{v}$\\}
@@
 
-\newcommand{\readCond}[1]{\plainfun{readCond}{#1}}
-\newcommand{\writeCond}[1]{\plainfun{writeCond}{#1}}
-\newcommand{\execCond}[1]{\plainfun{execCond}{#1}}
-\newcommand{\entryCond}[1]{\plainfun{enterCond}{#1}}
+\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
+\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
+\newcommand{\execCond}[1]{\plainfun{execCondition}{#1}}
+\newcommand{\entryCond}[1]{\plainfun{enterCondition}{#1}}
 
@@
   capabilities, and its operational semantics
-  (\sectionname~\ref{sec:capab-mach-with}).
+  (Section~\ref{sec:capab-mach-with}).
 \item We define a novel calling convention enforcing control-flow correctness and
-  encapsulation of stack frames (\sectionname~\ref{sec:stack-and-return-pointer}). It
+  encapsulation of stack frames (Section~\ref{sec:stack-and-return-pointer}). It
   relies solely on local capabilities and does not require OS support (like a
@@
   fundamental theorem expressing capability safety
-  are presented in \sectionname~\ref{sec:logical-relation}.
+  are presented in Section~\ref{sec:logical-relation}.
 \item We demonstrate our results by applying them to challenging examples,
@@
   control-flow correctness guarantees in the presence of cross-component
-  function pointers (\sectionname~\ref{sec:examples}). The examples demonstrate both
+  function pointers (Section~\ref{sec:examples}). The examples demonstrate both
   the power of our formulation of capability safety and our calling convention.
@@
 requirement is the most problematic aspect of our calling convention for
-performance, but see \sectionname~\ref{sec:discussion} for how this might be
+performance, but see Section~\ref{sec:discussion} for how this might be
 mitigated.
@@
 % \lau{If we need space, then maybe one of the examples is sufficient.}
-% \lau{Both of these are good to have their as the two last examples of \sectionname~\ref{sec:examples} use them.}
+% \lau{Both of these are good to have their as the two last examples of Section~\ref{sec:examples} use them.}
 %
@@
 \textbf{Calling convention} With the security measures introduced and motivated,
-let us summarize our proposed calling convention: \textit{At program start-up} A
-local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local
-capabilities. \textit{Before returning to the adversary} Clear non-return-value
-registers. Clear the part of the stack we had access to (not just the part we
-used). \textit{Before invoking the adversary} Push activation record to the
-stack. Create return pointer as local $\entry$-capability to the instructions in
-the record.  Restrict the stack capability to the unused part and clear
-it. Clear non-argument registers. \textit{Before invoking an adversary callback}
-Make sure callback is global. \textit{When invoked by an adversary} Make sure
-received stack pointer has permission $\rwlx$.
+let us summarize our proposed calling convention:
+\begin{description}[font=\normalfont\itshape]
+\item[At program start-up] A local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local capabilities.
+\item[Before returning to the adversary] Clear non-return-value registers. Clear
+  the part of the stack we had access to (not just the part we used).
+\item[Before invoking the adversary] Push activation record to the stack. Create
+  return pointer as local $\entry$-capability to the instructions in the record.
+  Restrict the stack capability to the unused part and clear it. Clear
+  non-argument registers.
+\item[Before invoking an adversary callback] Make sure callback is global.
+\item[When invoked by an adversary] Make sure received stack pointer has permission $\rwlx$.
+\end{description}
 
@@
 worlds that has been previously used in the literature (see
-\sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
+Section~\ref{sec:discussion}). However, temporary regions are new in our setting
 and there is an interesting interplay with the recursiveness of the worlds: for
@@
 \subsection{Worlds}
-\begin{wrapfigure}{R}{0.5\textwidth}
+\begin{figure}[tbp]
   \centering
@@
 
-  \caption{The relation between local/global capabilities and temporary/permanent
-    regions. The colored fields are regions governing parts of memory. Global capabilities cannot depend on temporary
+  \caption{An illustration of local/global capabilities and temporary/permanent
+    regions. The table represents a memory and the colored fields are regions
+    governing a memory range. Global capabilities cannot depend on temporary
     regions.}
   \label{fig:cap-world}
-\end{wrapfigure}
+\end{figure}
 % \dominique{Perhaps we could first try to convey the intuition of a permanent
@@
 % world-dependent. 
-As described in \sectionname~\ref{sec:formalizing-guarantees}, regions define sets of
+As described in Section~\ref{sec:formalizing-guarantees}, regions define sets of
 safe memory segments, but this set may itself be world-dependent. In other
@@
     \end{align*}
-  and for $W, W' \in \Wor\ldotp \quad  
-  \begin{aligned}
+  and for $W, W' \in \Wor$
+  \begin{align*}
     W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
     W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
-  \end{aligned}$
+  \end{align*}
 \end{theorem}
@@
 The \emph{private future world} relation $W' \futurestr W$ is defined similarly,
-using a private future region relation. The \emph{public future} region relation is the simplest. It satisfies
+using a private future region relation.
+
+The \emph{public future} region relation is the simplest. It satisfies
 % \dominique{is defined by?} 
@@
 public future world relations allows us to reinstate a region that has been revoked
-earlier. The \emph{private future region} relation satisfies:
+earlier.
+
+The \emph{private future region} relation satisfies:
 \begin{mathpar}
@@
     \text{ iff }\quad \left\{\begin{aligned}
-        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.35cm]
+        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.2cm]
         &\quad \forall r \in \activeReg{W} \ldotp\\
@@
         \union \\
-        % & \left\{
-        %     \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-        %   \right\}
-        % \union \\
+        & \left\{
+            \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+          \right\}
+        \union \\
         & \left\{
@@
           \right\} \union \\
-        % & \left\{
-        %     \npair{\stdcap[(\readwritel,\gl)] } \;\middle| \;
-        %     \begin{array}{l}
-        %      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-        %      \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
-        %     \end{array}
-        %    \right\}
-        % \union \\
-        % & \left\{
-        %   \npair{\stdcap[(\exec,\gl)]} \; \middle| \;
-        %   \begin{array}{l}
-        %      \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-        %      \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-        %   \end{array}
-        %    \right\}
-        % \union \\
+        & \left\{
+            \npair{\stdcap[(\readwritel,\gl)] } \;\middle| \;
+            \begin{array}{l}
+             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+             \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
+            \end{array}
+           \right\}
+        \union \\
+        & \left\{
+          \npair{\stdcap[(\exec,\gl)]} \; \middle| \;
+          \begin{array}{l}
+             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+             \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{array}
+           \right\}
+        \union \\
         & \left\{
@@
         \union \\
-        % & \left\{
-        %   \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
-        %   \begin{array}{l}
-        %     \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-        %     \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
-        %     \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-        %   \end{array}
-        %   \right\}
-        % \union \\
+        & \left\{
+          \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
+          \begin{array}{l}
+            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+            \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
+            \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{array}
+          \right\}
+        \union \\
         & \left\{
@@
              \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
-             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \\
+             \hfill \in \execCond{}(\gl)(W)
             \end{array}
-          \right\} \\
-          \union
-        & \dots \small{\text{\itshape and so on for permissions $\readonly$, $\readwritel$, $\exec$, and $\rwx$.}}
+          \right\}
       \end{aligned}
@@
     \right\} \\
-   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
+  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
+  & \quad\;\; \text{and } \gl = \glob \Rightarrow \future = \futurestr
   \end{align*}
@@
 configurations are safe according to the intuitions described in
-\sectionname~\ref{sec:formalizing-guarantees}. The definition is found in
+Section~\ref{sec:formalizing-guarantees}. The definition is found in
 \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
@@
 global. Note finally that the step-indexing described in
-\sectionname~\ref{sec:formalizing-guarantees} is effectively at play here, but hidden
+Section~\ref{sec:formalizing-guarantees} is effectively at play here, but hidden
 in the definition of the standard region $\iota^\pwl_{\start,\addrend}$, which
@@
 that they prohibit global write-local capabilities which, as discussed in
-\sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
+Section~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
 local capabilities.
@@
 of the values it has access to. We demonstrate this in
-\sectionname~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
+Section~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
 capabilities that point to arbitrary instructions. The relation between
 Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow
-correctness, will also be shown by example in \sectionname~\ref{sec:examples} as the
+correctness, will also be shown by example in Section~\ref{sec:examples} as the
 examples depend on these properties for correctness.
@@
 correctness properties for challenging program examples. The security
-measures of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to
+measures of Section~\ref{sec:stack-and-return-pointer} are deployed to
 ensure these properties. Since we are dealing with assembly language,
@@
 top of its stack and the assertion would fail. However, the security measures in
-\sectionname~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
+Section~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
 the return pointer is local, so it can only be stored on the stack, but the part
@@
      prepstk $r_\stk$
-     ${\text{\tiny\itshape\text{(continues in next column)}}}$
+     ${\tiny\text{\itshape(continues in next column)}}$
 \end{lstlisting}
@@
 \begin{lstlisting}
-     ${\text{\tiny\itshape\text{(continued from previous column)}}}$
+     ${\tiny\text{\itshape(continued from previous column)}}$
      store $x$ 0
@@
 adversary and an adversary callback invoked by us. As explained in
-\sectionname~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
+Section~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
 that the stack pointer that the closure receives from the adversary has
@@
   and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
-  Then if $
-  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
+  Then if
+  \[
+  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),
+  \]
   then $m'(\flag) = 0$.
@@
 
-As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
+As explained in Section~\ref{sec:stack-and-return-pointer}, the
 macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
@@
 \texttt{\footnotesize{f4}} returns.
-By utilizing the reinstation mechanism of the future world relation as well as our knowledge of the future worlds used, we can construct a world which the adversary's invariants are preserved.
-The details of this and the proofs of the other lemmas are found in the technical appendix\cite{technical_appendix}.
-
-\begin{comment}
+
 In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
@@
 \end{figure}
-\end{comment}
 \begin{comment}
@@
 security measures of the calling convention described in
-\sectionname~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
+Section~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
 calling convention without security guarantees. However, most of our security
@@
 \emph{Semantic, but not syntactic properties} The logical relation defined in
-\sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
+Section~\ref{sec:logical-relation} allows us to reason about capability machine
 programs. A limitation w.r.t. previous work is that the logical relation is
@@
 Finally, we summarize how our work relates to previous work. We do not
-repeat the work we discussed in \sectionname~\ref{sec:discussion}.
+repeat the work we discussed in Section~\ref{sec:discussion}.
 
@@
 of previous work. The capability machine formalized in
-\sectionname~\ref{sec:capab-mach-with} is a simple but representative
+Section~\ref{sec:capab-mach-with} is a simple but representative
 model, modeled mainly after the
@@
 will follow a legal path in the control-flow graph. As discussed in
-\sectionname~\ref{sec:discussion}, such a property ignores temporal properties and
+Section~\ref{sec:discussion}, such a property ignores temporal properties and
 seems hard to use for reasoning.
@@
 systems \cite{Ahmed:popl09}, public/private future worlds \cite{Dreyer:jfp12}
-(see \sectionname~\ref{sec:discussion} for a discussion), and biorthogonality
+(see Section~\ref{sec:discussion} for a discussion), and biorthogonality
 \cite{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

### 2017-09-29 esop-2018/paper.tex

--- 2017-09-29 esop-2018/paper.tex
@@
      prepstk $r_\stk$
-     ${\text{\tiny\itshape\text{(continues in next column)}}}$
+     ${\text{\tiny\itshape(continues in next column)}}$
 \end{lstlisting}
@@
 \begin{lstlisting}
-     ${\text{\tiny\itshape\text{(continued from previous column)}}}$
+     ${\text{\tiny\itshape(continued from previous column)}}$
      store $x$ 0

### 2017-09-29 fully-abstract-comp/technical-report.tex

@@
 If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
+\lau{28-09-2017: For now, the jump to a return pointer does the stack end point check. We could possibly move this to the proposed semantic condition as well.}
 \begin{align*}
@@
 \item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
-\item The base address of the stack capability is a fixed address, so in the semantice, it will be expressed as a constant that is publicly known.
+\item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
+\item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
+\item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
 \end{itemize}

### 2017-09-29 esop-2018/paper.tex
Attempt to make the paragraph about readCond more readable.

@@
 For capabilities with read permission, the $\readCond{}$ ensures that it can
-only be used to read safe words, i.e.\ words in the value relation.
-This is done by imposing an upper bound on the memory requirements of the region
-governing the addressed memory. The upper bound is formulated in terms of a
-standard region $\iota^\pwl_{\start,\addrend}$, which requires all the words in
-the range $[\start,\addrend]$ to be safe, i.e.\ in the value relation. The
-region governing the addressed memory must be at least as restrictive as the
-standard region. The read condition also takes into account the locality of the
-capability because, generally speaking, global capabilities should only depend
-on permanent regions. Concretely, we use the function
-$\var{localityReg}(\gl,W)$, which projects out all active (non-revoked) regions
-when the locality $\gl$ is local, but only the permanent regions when $\gl$ is
-global. Note finally that the step-indexing described in
-\sectionname~\ref{sec:formalizing-guarantees} is effectively at play here, but hidden
-in the definition of the standard region $\iota^\pwl_{\start,\addrend}$, which
-we omit for space reasons.
+only be used to read safe words, i.e.\ words in the value relation. To guarantee
+this, we require that the addressed memory is governed by a region $W(r)$ that
+imposes safety as a requirement on the values contained. This safety requirement
+is formulated in terms of a standard region $\iota^\pwl_{\start,\addrend}$. The
+definition of that standard island is omitted for space reasons, but it simply
+requires all the words in the range $[\start,\addrend]$ to be safe, i.e.\ in the
+value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{\start,\addrend}$
+means that $W(r)$ must accept only safe values like
+$\iota^\pwl_{\start,\addrend}$, but can be even more restrictive if desired. The
+read condition also takes into account the locality of the capability because,
+generally speaking, global capabilities should only depend on permanent regions.
+Concretely, we use the function $\var{localityReg}(\gl,W)$, which projects out
+all active (non-revoked) regions when the locality $\gl$ is local, but only the
+permanent regions when $\gl$ is global. Note finally that the step-indexing
+described in \sectionname~\ref{sec:formalizing-guarantees} is effectively at
+play here, but hidden in the definition of the standard region
+$\iota^\pwl_{\start,\addrend}$, which we omit for space reasons.
 

### 2017-09-29 esop-2018/paper.tex
terminology

@@
 is formulated in terms of a standard region $\iota^\pwl_{\start,\addrend}$. The
-definition of that standard island is omitted for space reasons, but it simply
+definition of that standard region is omitted for space reasons, but it simply
 requires all the words in the range $[\start,\addrend]$ to be safe, i.e.\ in the

### 2017-10-03 fully-abstract-comp/technical-report.tex
typo
Added a short section on related work and a bibtex file

@@
 \usepackage{stmaryrd}
+
+% References
+\usepackage{natbib}
+\bibliographystyle{plainnat}
 
@@
   \item $\ms$ and $[c_\code,c_\data]$ are self-contained with $\ms_\link$ as slack.
-  \item $\ms(\baddr_\code) = \seal(\normal,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr)$ where these are somehow the sufficient amount of seals.
+  \item $\ms(\baddr_\code) = \seal{\normal,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ where these are somehow the sufficient amount of seals.
   \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
@@
 
+\section{Related Work}
+\subsection{Conditional Full-Abstraction}
+The idea of conditional full-abstraction was used by \citet{Juglaret2016} to define full abstraction for unsafe languages.
+Their definition requires both the programs and the context to be fully defined (i.e.\ not cause undefined behavior).
+If the programs are not required to be fully-defined, then anything can happen which makes it impossible to reason about..
+In our work, undefined behavior marks cases that we do not want to consider because they should be excluded further up in the compilation chain.
+Further, if we have to take these cases into account, then we need to add checks which protects the trusted code against itself, but properly compiled code should not have to protect itself against itself.
+\lau{03-10-2017: We can adjust this when we have actually done something.}
+
+\bibliography{references}
 \end{document}

### 2017-10-05 fully-abstract-comp/technical-report.tex
Pre meeting commit.
New definitions related to programs and contexts.

@@
 \usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{mathrsfs}
 \usepackage{thmtools}
@@
 \newcommand{\totherwise}{\text{otherwise }}
+
+% Equivalences
+\newcommand{\sconeq}{\mathrel{\src{\approx_{\mathrm{ctx}}}}}
+\newcommand{\tconeq}{\mathrel{\approx_{\mathrm{ctx}}}}
 
@@
 \newcommand{\tRegName}{\targetdom{RegisterName}}
+
+%%% Programs and contexts
+\newcommand{\program}{\mathscr{P}}
+\newcommand{\context}{\mathscr{C}}
+
+\newcommand{\plug}[2]{#1[#2]}
+
+%%% Operational semantics
+\newcommand{\step}{\rightarrow}
+\newcommand{\nstep}[1][n]{\step_{#1}}
+\newcommand{\steps}{\step^*}
+\newcommand{\diverge}{\Uparrow}
 
@@
 \newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
+\newcommand{\encInstr}[1]{\plainfun{encodeInstruction}{#1}}
+\newcommand{\decInstr}[1]{\plainfun{decodeInstruction}{#1}}
 \newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
@@
 
+\subsubsection{Step Relation}
+
 \subsubsection{Instruction Interpretation}
@@
   \begin{itemize}
-  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$
+  \item We say that a memory segment $\ms$ and a list of words $\bar{w}$
     are self-contained iff
   \[
-    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom{\ms}
+    \forall w \in \bar{w} \ldotp w \in \Caps \lor w = \stkptr{\_,\_,\_,\_} \Rightarrow \range{w} \subseteq \dom(\ms)
   \]
@@
 
+\begin{definition}
+  \label{def:wxorx}
+  ~\\
+  A word $w$ is said to respect write-xor-execute iff
+  \begin{itemize}
+  \item $w \in \ints$
+  \item $w = \seal{\_,\_,\_,\_}$
+  \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
+  \item $\seal{\_,\vsc}$ and $\vsc$ respects write-xor-execute
+  \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
+  \item $\src{w = \retptrc{\_}}$
+  \item $\src{w = \retptrd{\_}}$
+  \end{itemize}
+~\\
+  A register file $\reg$ is said to respect write-xor-execute with slack $[\bar{r}]$ iff \\
+  For all $r \in RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
+~\\
+~\\
+  A memory segment $\ms$ is said to respect write-xor-execute iff\\
+  For all $\aaddr \in \dom(\ms)$ $\ms(\aaddr)$ respects write-xor-execute
+\end{definition}
+
+\begin{definition}
+  \label{def:no-stk-ptr}
+~\\
+  We say that a word $w$ covers no stack pointer iff\\
+  \begin{item}
+    \item $w \neq ((\perm,\linear),\stkb,\_,\_)$\\
+          $\src{w \neq \stkptr{\_,\_,\_,\_}}$
+    \item if $w = \sealed{\_,\vsc}$, then $\vsc$ covers no stack pointer.
+  \end{item}
+~\\
+  We say that a memory segment $\ms$ contains no stack pointer iff
+  \[
+    \forall \aaddr \in \dom(\ms)\ldotp \text{$\ms(\aaddr)$ covers no stack pointer}
+  \]
+~\\
+  We say that a register file $\reg$ contains no stack pointer with slack $[\bar{r}]$ iff
+  \[
+    \forall r \in \reg \setminus [\bar{r}] \ldotp \text{$\reg(r)$ covers no stack pointer}
+  \]
+\end{definition}
+
+\begin{definition}
+  \label{def:resp-linearity}
+~\\
+  We say that a memory segment $\ms$ and words $\bar{w}$ respects linearity iff\\
+  Let $\var{lc}$ be all the linear capabilities in $\ms$ and $\bar{w}$, i.e.,
+  \[
+    \var{lc} = \left\{c \middle|
+      \begin{array}{l}
+        \exists \aaddr \in \dom(\ms) \ldotp c = \ms(\aaddr) = ((\_,\linear),\_,\_,\_) \\
+        \exists \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) = \sealed{\_,c} \land c = ((\_,\linear),\_,\_,\_) \\
+        \src{\exists \aaddr \in \dom(\ms) \ldotp c = \ms(\aaddr) = \stkptr{\_,\_,\_,\_}} \\
+        \src{\exists \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) = \sealed{\_,c} \land c= \stkptr{\_,\_,\_,\_}} \\
+      \end{array}
+\right\}
+  \]
+\end{definition}
+
 \subsubsection{Programs}
 Here we define what constitutes a program.
-
+\lau{05-10-2017: It seems like both program and context takes care of linking atm.}
 \begin{definition}
@@
   \lau{27-09-2017: The ``somehow''s need to be fixed.}
-  A program consists of two sealed capabilities, $\vsc_\code = \sealed{c_\code,\sigma}$ and $\vsc_\data = \sealed{c_\data,\sigma}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
-  \begin{itemize}
-  \item $c_\code = ((\rx,\lin_\code),\baddr_\code,\eaddr_\code,\aaddr_\code)$ where $\aaddr = \baddr + 3$ and $[\baddr_\code,\eaddr_\code]\subseteq \dom(\ms)$
+  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{c_\code,\sigma}$ and $\vsc_\data = \sealed{c_\data,\sigma}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
+  \begin{itemize}
+  \item $c_\code = ((\rx,\lin_\code),\baddr_\code,\eaddr_\code,\aaddr_\code)$ where $\aaddr = \baddr + 2$ and $[\baddr_\code,\eaddr_\code]\subseteq \dom(\ms)$
   \item $c_\data = ((\rw,\lin_\data),\baddr_\data,\eaddr_\data,\baddr_\data)$ where $[\baddr_\data,\eaddr_\data] \subseteq \dom(\ms)$ and $[\baddr_\code,\eaddr_\code] \cap [\baddr_\data,\eaddr_\data] = \emptyset$
   \item $\ms$ and $[c_\code,c_\data]$ are self-contained with $\ms_\link$ as slack.
+  \item $\ms$ respects write-xor-execute.
   \item $\ms(\baddr_\code) = \seal{\normal,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ where these are somehow the sufficient amount of seals.
   \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
-  \item $\ms(\baddr_\code+2) = \dots$ where $\dots$ is somehow the stack height.
   \end{itemize}
@@
 \subsubsection{Contexts}
+\begin{definition}
+  \label{def:context}
+  A context, $\context$, is a register file $\reg$, registers $r_1$ and $r_2$, and memory segments $\ms$ and $\ms_\link$ satisfying
+  \begin{itemize}
+  \item $\reg$ respects write-xor-execute with slack $[r_1,r_2]$.
+  \item $\reg(\rstk) = ((\rw,\linear),\stkb,\eaddr,\eaddr)$ where $\stkb \leq \eaddr$\\
+        {\sourcecolor{}$\reg(\rstk) = \stkptr{\rw,\stkb,\eaddr,\eaddr}$ where $\stkb \leq \eaddr$}
+  \item $\ms \uplus \ms_\link$ respects write-xor-execute
+  \item $\ms \uplus \ms_\link$ and $[\reg(\RegName)]$ are self-contained.
+  \item $\ms \uplus \ms_\link$ contains no stack pointer.
+  \item $\reg$ contains no stack pointer with slack $[r_1,r_2]$.
+  \item {\color{red}TODO: No return pointer}
+  \end{itemize}
+In this definition, the source level specific conditions replace the equivalent target level conditions.
+\end{definition}
+
+\subsubsection{Plugging a context}
 
@@
 \clearpage
+\section{Full Abstraction}
+\begin{definition}
+  \label{def:points-to-instr}
+  A configuration $(\ms,\reg)$ {\sourcecolor or $(\ms,\reg,\_,\_)$} os said to point instruction $i$ in $A$ iff
+  \[
+    \reg(\pcreg) = ((\_,\_),\_,\_,\aaddr) \tand \aaddr \in A
+  \]
+and
+\[
+  \decInstr{\ms(\aaddr)} = i
+\]
+\end{definition}
+
+We want to make a conditional full-abstraction statement.
+The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do.
+(2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation
+\begin{definition}
+  \label{def:check-stack-addr-before-call}
+  A configuration $\src{\Phi}$ is said to have reasonable calls in $A$ for $n$ steps iff\\
+  for all $i \leq n$ if
+  \[
+    \src{\Phi} \nstep[i] \src{\Phi'}
+  \]
+  and $\src{\Phi'}$ points to $\src{\scall{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2'}$ and there exists no $j<i$ s.t.
+  \[
+    \src{\Phi} \nstep[j] \src{\Phi''}
+  \]
+  where $\src{\Phi''}$ points to $\src{\scall{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
+  then
+  \[
+    \src{\Phi}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
+  \]
+  and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
+\end{definition}
+
+\begin{definition}
+  A program $\program$ with memory segment $\ms_\program$ is said to be reasonably compiled if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  \begin{itemize}
+  \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
+  \end{itemize}
+\end{definition}
+
+\begin{theorem}
+  \label{thm:full-abstraction}
+  For programs $\program_1$ and $\program_2$ that are reasonably compiled compiled, we have
+  \begin{gather*}
+    \src{\program_1} \sconeq \src{\program_2}\\
+    \Updownarrow\\
+    \comp{\src{\program_1}} \tconeq \comp{\src{\program_2}}
+  \end{gather*}
+\end{theorem}
+
+\clearpage
 \section{Examples}

### 2017-10-09 esop-2018/paper.tex
Attempt to further emphasize the points we made in our POPL author response.

@@
   provide a formalization of a representative capability machine with local
-  capabilities and propose a novel calling convention for enforcing control-flow
+  capabilities and study a novel calling convention for enforcing control-flow
   correctness and encapsulation of local state on it. To prove these properties,
@@
 invocation whereafter the capability can be revoked. However, details are scarce
-and a lot remains to be investigated (e.g., how to combine disjoint stacks with
-cross-domain function pointers in a scalable way?) and there is no argument that
-the approach is watertight.
+(how does it work precisely? what features are supported?) and a lot remains to
+be investigated (e.g., combining disjoint stacks with cross-domain function
+pointers seems like it will scale poorly to large numbers of components?).
+Finally, there is no argument that the approach is watertight and it is not even
+clear what security property is targeted exactly.
 
@@
   our calling convention or a specific way of using the system's capabilities.
+  We are the first to apply these techniques for reasoning about capability machines and we believe they will prove useful for many other purposes than our calling convention.
 \item We introduce two novel technical ideas in the unary, step-indexed Kripke

### 2017-10-09 esop-2018/paper.tex
Shorten a contributions paragraph and suggest some more text to be cut in related work...

@@
 \item We introduce two novel technical ideas in the unary, step-indexed Kripke
-  logical relation used to formulate the above theorem. First, we model the
-  natural continuation-passing-style of assembly code (with return pointers as
-  continuations) using a \emph{single} orthogonal closure, rather than
-  the earlier used biorthogonal closure. Second, we express the special nature of
-  local capabilities using a variant of Dreyer et. al.'s public and private
-  future worlds~\cite{Dreyer:jfp12}, which they used to express well-bracketedness
-  guarantees in a lambda calculus. The logical relation and the
-  fundamental theorem expressing capability safety
-  are presented in \sectionname~\ref{sec:logical-relation}.
+  logical relation used to formulate the above theorem: the use of a
+  \emph{single} orthogonal closure (rather than the earlier used biorthogonal
+  closure) and a variant of Dreyer et. al.'s public and private future
+  worlds~\cite{Dreyer:jfp12} to express the special nature of local
+  capabilities. The logical relation and the fundamental theorem expressing
+  capability safety are presented in \sectionname~\ref{sec:logical-relation}.
 \item We demonstrate our results by applying them to challenging examples,
@@
 capability safety for a lambda-calculus (with concurrency) compared to the
-earlier work by \cite{Devriese:2016ObjCap}. In the future we would also like to
+earlier work by \cite{Devriese:2016ObjCap}.
+%domi: rest of the paragraph could be cut?
+In the future we would also like to
 investigate a new program logic for reasoning about capability safety for our

### 2017-10-10 esop-2018/paper.tex
Rephrase the contributions paragraph.

@@
 
-In this paper, we make progress towards the goal of efficient and provably
-watertight enforcement of encapsulation and control-flow correctness on
-capability machines. Specifically, we make the following contributions:
+In this paper, we make two main contributions: (1) an alternative calling
+convention that uses local capabilities to enforce stack frame encapsulation and
+well-bracketed control flow, and (2) perhaps more importantly, we adapt and
+apply the well-studied techniques of step-indexed Kripke logical relations for
+reasoning about code on a representative capability machine code with local
+capabilities in general and correctness and security of the calling convention
+in particular. More specifically, we make the following contributions:
 \begin{itemize}

### 2017-10-10 esop-2018/paper.tex
sec. 4

@@
 \label{sec:logical-relation}
-Now that we have defined our calling convention, how can we be sure that it
-works? More concretely, suppose that we have a program that uses the convention
-in its interaction with untrusted adversary code. Can we formally prove the
-program's correctness if it relies on well-bracketed control flow and private
-state encapsulation for the interaction with the adversary? Clearly, such a
-proof should depend on a formal expression of the guarantees provided by the
-capability machine, including the specific guarantees for local capabilities.
-
-In this section, we construct such a formalization. We make use of some
-well-studied and powerful (but non-trivial) machinery from the literature.
-Specifically, we employ a unary step-indexed Kripke logical relation with
-recursive worlds, and some additional special characteristics of our own.
-Step-indexing, Kripke logical relations and recursive worlds are techniques that
-may be familiar from lambda calculus settings, but it may not be clear to the
-reader how they apply in this more low-level assembly language. Therefore, we do
-not immediately dive into the details, but we first try to provide some informal
-intuition about how all of this machinery comes into play in our setting, in the
-next section.
-
-Note: even though the calling convention is the main application in this paper,
-the logical relation we construct is very general and should be regarded as an
-independent contribution.
-
-\subsection{Formalizing the guarantees of the capability machine}
-\label{sec:formalizing-guarantees}
-What differentiates a capability machine from a more standard assembly language
-is that we can bound the authority of an executing block of code, based solely
-on the capabilities it has access to. Specifically, it does not matter which
-instructions are actually executed, i.e., the bound also applies to untrusted
-adversary code that has not been inspected or modified in any way.
-
-\paragraph{Worlds}
-But what does a ``bound on the authority'' of an executing block of code mean?
-In our setting, there are no externally observable side-effects and the only
-primitive authority that code may hold is authority over memory. As such, the
-authority bounds we consider are related to memory, but in a form that is more
-fine-grained than standard read/write authority: a piece of code's authority can be bounded
-by arbitrary memory invariants that it is required to respect. Specifically, we
-will define worlds $W \in \Worlds$, which describe a set of memory invariants, and
-our results will express authority bounds on code as \emph{safety with respect to such
-a world,} i.e., the fact that the code respects the invariants registered in
-the world.
-
-\paragraph{Safe values}
-So, let's say that we have a world $W$ expressing that the memory must contain
-value $42$ at address $0$, may contain arbitrary values at
-addresses $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$, and
-an integer at address $100$ that may only increase over time\footnote{Indeed, we
-  will allow a notion of \emph{evolvable} invariants, aka \emph{protocols}, that can
-  express such a temporal property.}. Our main theorem will state that if the
-current register file only contains safe words (numbers or capabilities which
-preserve the invariants in $W$ under any interaction), then an execution will
-necessarily also preserve the memory invariants (irrespective of the instructions being
-executed).
-
-To make this more precise, we need to define the set $\stdvr(W) \in
-\powerset{\Words}$ of words that are safe w.r.t.\ $W$. Essentially, the set
-should only include words that preserve $W$'s invariants under any interaction,
-but should otherwise be as liberal as possible. Numbers are clearly always safe,
-as they cannot be used to break invariants. Whether a capability is safe depends
-on the authority that it carries. In the above-described world, a read capability
-for address $0$ is safe, as it can only be used to read the value $42$, which is
-itself safe. However, a write capability for address $0$ is not safe: it can be
-used to overwrite the memory at that address with a value other than $42$,
-breaking the invariant for that address.
-
-\paragraph{Step-indexing}
-More generally, we want to define that a read capability for memory range
-$[b,e]$ is safe, if the world guarantees that the words at those addresses are
-themselves safe. However, this definition is cyclic: what if the world
-guarantees that the memory at address $a$ will contain a read capability for
-address $a$? The definition then just says that a read capability for address
-$a$ is safe if and only if the same read capability for address $a$ is safe. This form of
-cyclic reasoning is related to similar challenges in languages with recursive
-types or higher-order ML-style references, and a standard solution is
-to use step-indexing \cite{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
-certain number of interaction steps. All words will be considered
-safe up to $0$ steps (since if there is no interaction, nothing unsafe
-can happen), and, for example, a read capability will be safe up to $n$
-steps if the world guarantees that the words at the corresponding addresses are
-safe up to $n - 1$ steps. We can then prove that the above read capability for
-address $a$ is safe up to any number of steps.
-
-\paragraph{Future worlds}
-So worlds are defined as a set of invariants on the memory, but what if we
-allocate fresh memory through malloc? We may want to establish new invariants
-for this freshly allocated memory and be sure that the adversary will also
-respect those (if we don't provide them with capabilities through which the new
-invariants can be broken). To accommodate this, we allow worlds to evolve, for
-example by adding additional invariants for freshly allocated memory. Formally,
-we define valid ways for a world $W$ to evolve into a new world $W'$ through a
-future-world relation $W' \future W$ and we ensure that the set of safe words in
-world $W$ must remain safe in any future world $W'$. Defining safety w.r.t.\ a
-notion of evolvable worlds makes our logical relation into a \emph{Kripke}
-logical relation \cite{pitts_operational_1998}.
-
-\paragraph{Invariants and Recursive Worlds}
-So worlds group a set of memory invariants, but how are they actually
-defined formally? We represent each invariant by a region
-$r \in \Regions$. We will see later that regions contain state
-machines to support a notion of evolvable invariants, but in every
-state, they also contain a predicate $H$ that defines which memory
-segments are acceptable in the current state of the invariant.
-Unfortunately, it is not enough to just take
-$H \in \powerset{\MemSegments}$, because sometimes the invariant may
-itself be world-dependent. For example, we may want to express
-invariants like ``the memory at address $50$ contains a value that is
-safe in the current world''. As explained, worlds may evolve, and the
-set of safe values may grow in future worlds, and therefore we need to
-index $H$ over worlds, i.e., take
-$H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
-with worlds containing regions with world-indexed predicates,
-i.e., the set of worlds must be recursively defined. We will see how such a recursive
-definition can be accommodated using techniques from the
-literature (essentially an advanced application of step-indexing).
-
-\paragraph{Local capabilities}
-A final thing we provide informal intuition for is how our results take into account local
-capabilities and their special treatment by the hardware. Normally, when we
-invoke an untrusted piece of code and provide it with certain global
-capabilities, it may have stored those capabilities in memory and we will only
-be able to reinvoke the code if we can guarantee that those values are still
-valid. Formally, worlds represent the invariants that global capabilities'
-safety relies on and the reinvocation is only safe in future worlds, where the
-invariants are respected.
-
-However, if we provide the adversary with local capabilities in that first
-invocation, then the situation is a bit different. The adversary has no way to
-store these local capabilities, so if we make sure that there are also no old
-local capabilities in the register file for the second invocation (including the
-capability being invoked), then the adversary cannot use them any more.%
-\footnote{We are ignoring write-local capabilities in this discussion. If the
-  adversary does have access to write-local capabilities in the first and second
-  invocation, then the memory they address must entirely be cleared before the
-  second invocation in order for the reinvocation to remain safe.} Therefore, we
-can allow the second invocation to happen in \emph{private} future worlds ($W'
-\futurestr W$), in which safe global capabilities remain safe, but local ones
-don't. This private future world relation is more liberal than the standard,
-public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
-safe). Concretely, worlds may contain \emph{temporary} regions, representing
-invariants that only local capabilities may rely on for their safety, and which
-may be revoked (disabled) in private future worlds.
-
-Interestingly, this idea is a variant of a notion of public/private future
-worlds that has been previously used in the literature (see
-\sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
-and there is an interesting interplay with the recursiveness of the worlds: for
-a temporary region, the predicate $H \in \Worlds \rightarrow
-\powerset{\MemSegments}$ (which defines the safe memory segments in the
-current world) is only required to be monotone w.r.t.\ public
-future worlds (i.e.\ safe memory segments remain safe in public future worlds),
-but for permanent regions, it must be monotone w.r.t.\ private future worlds.
-As a consequence, the memory for a permanent region may not contain local
-capabilities (as their safety would be broken in private future worlds), which
-in turn implies that only local capabilities may have write-local permission (a
-general sanity requirement when using local capabilities\footnote{In fact, local
-  capabilities become useless as soon as the adversary has access to a single
-  global, write-local capability.})\footnote{For explanation purposes, this
-  discussion ignores certain ways to allow for local capabilities in a permanent
-  region, for example, by not requiring that they are valid or requiring that
-  they are local versions of valid global capabilities.}.
+
+In this section, we formalize the guarantees provided by the
+capability machine, including the specific guarantees for local
+capabilities, by means of a step-indexed Kripke logical relation 
+with recursively defined worlds. 
+We use the logical relation in the following section to show
+local-state encapsulation and control-flow integrity properties for
+challenging example programs.
+
+% Our logical relation will prescribe which values are capability safe given a
+% world $W$ containing a set of memory invariants. For expressiveness,
+% we use invariants which can evolve over time (sometimes also referred
+% to as \emph{protocols}). We thus start by describing the worlds we
+% will use.
+
+% Now that we have defined our calling convention, how can we be sure that it
+% works? More concretely, suppose that we have a program that uses the convention
+% in its interaction with untrusted adversary code. Can we formally prove the
+% program's correctness if it relies on well-bracketed control flow and private
+% state encapsulation for the interaction with the adversary? Clearly, such a
+% proof should depend on a formal expression of the guarantees provided by the
+% capability machine, including the specific guarantees for local capabilities.
+
+% In this section, we construct such a formalization. We make use of some
+% well-studied and powerful (but non-trivial) machinery from the literature.
+% Specifically, we employ a unary step-indexed Kripke logical relation with
+% recursive worlds, and some additional special characteristics of our own.
+% Step-indexing, Kripke logical relations and recursive worlds are techniques that
+% may be familiar from lambda calculus settings, but it may not be clear to the
+% reader how they apply in this more low-level assembly language. Therefore, we do
+% not immediately dive into the details, but we first try to provide some informal
+% intuition about how all of this machinery comes into play in our setting, in the
+% next section.
+
+% Note: even though the calling convention is the main application in this paper,
+% the logical relation we construct is very general and should be regarded as an
+% independent contribution.
+
+% \subsection{Formalizing the guarantees of the capability machine}
+% \label{sec:formalizing-guarantees}
+% What differentiates a capability machine from a more standard assembly language
+% is that we can bound the authority of an executing block of code, based solely
+% on the capabilities it has access to. Specifically, it does not matter which
+% instructions are actually executed, i.e., the bound also applies to untrusted
+% adversary code that has not been inspected or modified in any way.
+
+% \paragraph{Worlds}
+% But what does a ``bound on the authority'' of an executing block of code mean?
+% In our setting, there are no externally observable side-effects and the only
+% primitive authority that code may hold is authority over memory. As such, the
+% authority bounds we consider are related to memory, but in a form that is more
+% fine-grained than standard read/write authority: a piece of code's authority can be bounded
+% by arbitrary memory invariants that it is required to respect. Specifically, we
+% will define worlds $W \in \Worlds$, which describe a set of memory invariants, and
+% our results will express authority bounds on code as \emph{safety with respect to such
+% a world,} i.e., the fact that the code respects the invariants registered in
+% the world.
+
+% \paragraph{Safe values}
+% So, let's say that we have a world $W$ expressing that the memory must contain
+% value $42$ at address $0$, may contain arbitrary values at
+% addresses $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$, and
+% an integer at address $100$ that may only increase over time\footnote{Indeed, we
+%   will allow a notion of \emph{evolvable} invariants, aka \emph{protocols}, that can
+%   express such a temporal property.}. Our main theorem will state that if the
+% current register file only contains safe words (numbers or capabilities which
+% preserve the invariants in $W$ under any interaction), then an execution will
+% necessarily also preserve the memory invariants (irrespective of the instructions being
+% executed).
+
+% To make this more precise, we need to define the set $\stdvr(W) \in
+% \powerset{\Words}$ of words that are safe w.r.t.\ $W$. Essentially, the set
+% should only include words that preserve $W$'s invariants under any interaction,
+% but should otherwise be as liberal as possible. Numbers are clearly always safe,
+% as they cannot be used to break invariants. Whether a capability is safe depends
+% on the authority that it carries. In the above-described world, a read capability
+% for address $0$ is safe, as it can only be used to read the value $42$, which is
+% itself safe. However, a write capability for address $0$ is not safe: it can be
+% used to overwrite the memory at that address with a value other than $42$,
+% breaking the invariant for that address.
+
+% \paragraph{Step-indexing}
+% More generally, we want to define that a read capability for memory range
+% $[b,e]$ is safe, if the world guarantees that the words at those addresses are
+% themselves safe. However, this definition is cyclic: what if the world
+% guarantees that the memory at address $a$ will contain a read capability for
+% address $a$? The definition then just says that a read capability for address
+% $a$ is safe if and only if the same read capability for address $a$ is safe. This form of
+% cyclic reasoning is related to similar challenges in languages with recursive
+% types or higher-order ML-style references, and a standard solution is
+% to use step-indexing \cite{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
+% certain number of interaction steps. All words will be considered
+% safe up to $0$ steps (since if there is no interaction, nothing unsafe
+% can happen), and, for example, a read capability will be safe up to $n$
+% steps if the world guarantees that the words at the corresponding addresses are
+% safe up to $n - 1$ steps. We can then prove that the above read capability for
+% address $a$ is safe up to any number of steps.
+
+% \paragraph{Future worlds}
+% So worlds are defined as a set of invariants on the memory, but what if we
+% allocate fresh memory through malloc? We may want to establish new invariants
+% for this freshly allocated memory and be sure that the adversary will also
+% respect those (if we don't provide them with capabilities through which the new
+% invariants can be broken). To accommodate this, we allow worlds to evolve, for
+% example by adding additional invariants for freshly allocated memory. Formally,
+% we define valid ways for a world $W$ to evolve into a new world $W'$ through a
+% future-world relation $W' \future W$ and we ensure that the set of safe words in
+% world $W$ must remain safe in any future world $W'$. Defining safety w.r.t.\ a
+% notion of evolvable worlds makes our logical relation into a \emph{Kripke}
+% logical relation \cite{pitts_operational_1998}.
+
+% \paragraph{Invariants and Recursive Worlds}
+% So worlds group a set of memory invariants, but how are they actually
+% defined formally? We represent each invariant by a region
+% $r \in \Regions$. We will see later that regions contain state
+% machines to support a notion of evolvable invariants, but in every
+% state, they also contain a predicate $H$ that defines which memory
+% segments are acceptable in the current state of the invariant.
+% Unfortunately, it is not enough to just take
+% $H \in \powerset{\MemSegments}$, because sometimes the invariant may
+% itself be world-dependent. For example, we may want to express
+% invariants like ``the memory at address $50$ contains a value that is
+% safe in the current world''. As explained, worlds may evolve, and the
+% set of safe values may grow in future worlds, and therefore we need to
+% index $H$ over worlds, i.e., take
+% $H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
+% with worlds containing regions with world-indexed predicates,
+% i.e., the set of worlds must be recursively defined. We will see how such a recursive
+% definition can be accommodated using techniques from the
+% literature (essentially an advanced application of step-indexing).
+
+% \paragraph{Local capabilities}
+% A final thing we provide informal intuition for is how our results take into account local
+% capabilities and their special treatment by the hardware. Normally, when we
+% invoke an untrusted piece of code and provide it with certain global
+% capabilities, it may have stored those capabilities in memory and we will only
+% be able to reinvoke the code if we can guarantee that those values are still
+% valid. Formally, worlds represent the invariants that global capabilities'
+% safety relies on and the reinvocation is only safe in future worlds, where the
+% invariants are respected.
+
+% However, if we provide the adversary with local capabilities in that first
+% invocation, then the situation is a bit different. The adversary has no way to
+% store these local capabilities, so if we make sure that there are also no old
+% local capabilities in the register file for the second invocation (including the
+% capability being invoked), then the adversary cannot use them any more.%
+% \footnote{We are ignoring write-local capabilities in this discussion. If the
+%   adversary does have access to write-local capabilities in the first and second
+%   invocation, then the memory they address must entirely be cleared before the
+%   second invocation in order for the reinvocation to remain safe.} Therefore, we
+% can allow the second invocation to happen in \emph{private} future worlds ($W'
+% \futurestr W$), in which safe global capabilities remain safe, but local ones
+% don't. This private future world relation is more liberal than the standard,
+% public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
+% safe). Concretely, worlds may contain \emph{temporary} regions, representing
+% invariants that only local capabilities may rely on for their safety, and which
+% may be revoked (disabled) in private future worlds.
+
+% Interestingly, this idea is a variant of a notion of public/private future
+% worlds that has been previously used in the literature (see
+% \sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
+% and there is an interesting interplay with the recursiveness of the worlds: for
+% a temporary region, the predicate $H \in \Worlds \rightarrow
+% \powerset{\MemSegments}$ (which defines the safe memory segments in the
+% current world) is only required to be monotone w.r.t.\ public
+% future worlds (i.e.\ safe memory segments remain safe in public future worlds),
+% but for permanent regions, it must be monotone w.r.t.\ private future worlds.
+% As a consequence, the memory for a permanent region may not contain local
+% capabilities (as their safety would be broken in private future worlds), which
+% in turn implies that only local capabilities may have write-local permission (a
+% general sanity requirement when using local capabilities\footnote{In fact, local
+%   capabilities become useless as soon as the adversary has access to a single
+%   global, write-local capability.})\footnote{For explanation purposes, this
+%   discussion ignores certain ways to allow for local capabilities in a permanent
+%   region, for example, by not requiring that they are valid or requiring that
+%   they are local versions of valid global capabilities.}.
 
@@
 
-As describe above, worlds are a collection of invariants. Formally, a world is a
+%As describe above, worlds are a collection of invariants. Formally,
+
+A world is a
 finite map from region names, modeled as natural numbers, to regions that each
@@
 % world-dependent. 
-As described in \sectionname~\ref{sec:formalizing-guarantees}, regions define sets of
-safe memory segments, but this set may itself be world-dependent. In other
-words, our worlds are defined recursively. Recursive worlds are a common
-requirement in Kripke models and the following lemma uses the method of
+%As described in \sectionname~\ref{sec:formalizing-guarantees}, regions define sets of
+
+Regions are used to define safe memory segments, but this set may itself be world-dependent. In other
+words, our worlds are defined recursively. Recursive worlds are common
+in Kripke models and the following lemma uses the method of
 \cite{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
@@
 protocols whereas temporary regions can rely on both temporary and permanent
-protocols.  $\UPred{\HeapSegments}$ is a step-indexed, downwards closed
-predicate on memory segments.
+protocols.  $\UPred{\HeapSegments}$ is the set of step-indexed, downwards closed
+predicates on memory segments:
+$\UPred{\HeapSegments} = \{ A \subseteq  \nats \times
+\HeapSegments \mid 
+\forall (n, ms) \in A. \forall m
+\leq n. (m, ms) \in A\}$. 
+\
 
@@
 \end{align*}
-Note that it is possible to state nonsensical worlds with conflicting
-requirements for the same part of memory. There would, however, be no
-memory that would satisfy such a world which for all intents and purposes would
-make it discounted by the logical relation.
+% LB: omitted the following for brevity:
+% Note that it is possible to state nonsensical worlds with conflicting
+% requirements for the same part of memory. There would, however, be no
+% memory that would satisfy such a world which for all intents and purposes would
+% make it discounted by the logical relation.
 
@@
 %   respect them itself too.} 
-The logical relation defines semantically when values, program counters, and
-configurations are safe according to the intuitions described in
-\sectionname~\ref{sec:formalizing-guarantees}. The definition is found in
-\figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and we provide some
-explanations in the following paragraphs. We omit some definitions and explain
-them only verbally, but precise definitions can be found in the technical
+The logical relation defines semantically when values, program
+counters, and configurations are capability safe. The definition is
+found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and
+we provide some explanations in the following paragraphs. For space
+reasons, we omit some definitions and explain them only verbally, but
+precise definitions can be found in the technical
 appendix~\cite{technical_appendix}.
@@
 First, the \emph{observation relation} $\observations$ defines what
-configurations we consider safe. As previously stated, a configuration is safe
-with respect to a world when the execution of said configuration does not break
+configurations we consider safe. A configuration is safe
+with respect to a world, when the execution of said configuration does not break
 the memory protocols of the world. Roughly speaking, this means that when the
@@
 
-The \emph{value relation} $\stdvr$ defines when words are safe, i.e.\ when they
-cannot be used to break the memory protocols defined by the world. We make the
+The \emph{value relation} $\stdvr$ defines when words are safe. We make the
 value relation as liberal as possible by considering what is the most we can
@@
 kind of permission a capability can have.
-
+{
 For capabilities with read permission, the $\readCond{}$ ensures that it can
@@
 all active (non-revoked) regions when the locality $\gl$ is local, but only the
-permanent regions when $\gl$ is global. Note finally that the step-indexing
-described in \sectionname~\ref{sec:formalizing-guarantees} is effectively at
-play here, but hidden in the definition of the standard region
-$\iota^\pwl_{\start,\addrend}$, which we omit for space reasons.
+permanent regions when $\gl$ is global. The 
+the definition of the standard region
+$\iota^\pwl_{\start,\addrend}$ can be found
+in~\cite{technical_appendix}; it makes use of the isomorphism from
+Theorem~\ref{thm: world-existence}.
 
@@
 safe word to be written there. In order to make the logical relation as liberal
-as possible, we make this a lower bound of what the region may allow. From
-another point of view, this lower bound is also an upper bound on the amount of
-trust that can be placed in the contents of memory that an adversary has been
-given write access to. For write capabilities, we also have to take into account
+as possible, we make this a lower bound of what the region may allow. 
+% LB: omitted the following for brevity:
+% From another point of view, this lower bound is also an upper bound on the amount of
+% trust that can be placed in the contents of memory that an adversary has been
+% given write access to. 
+For write capabilities, we also have to take into account
 the two flavours of write permissions: write and write-local. In the case of
@@
 
-In the technical appendix, we prove that (as explained before), safety of all
-values is preserved in public future worlds, but safety of global values is
+In the technical appendix, we prove that safety of all values is
+preserved in public future worlds, and that safety of global values is
 also preserved in private future worlds:
@@
 the capability machine. Essentially, it says a capability that only grants safe
-authority is capability safe as a program counter (where safety, as in the
-previous sections, means respect of memory protocols).
+authority is capability safe as a program counter.
 \begin{theorem}[Fundamental Theorem]
@@
 examples depend on these properties for correctness.
-
-From another point of view, the Fundamental Theorem also serves as a sanity
-check of the logical relation as it guarantees a large class of
-capabilities actually inhabits it.
+See the technical appendix~\cite{technical_appendix} for a detailed
+proof (by induction over the step-index $n$) of the theorem.
+% LB: following omitted for brevity:
+% From another point of view, the Fundamental Theorem also serves as a sanity
+% check of the logical relation as it guarantees a large class of
+% capabilities actually inhabits it.
 
@@
 
-\cite{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
+Swasey et. al.~\cite{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
 of object capability patterns. The logic is based on
@@
 %domi: rest of the paragraph could be cut?
-In the future we would also like to
-investigate a new program logic for reasoning about capability safety for our
-capability machine model. We think Iris would also be a natural starting point
-for such an endeavour, since Iris is really a framework, which can be
-instantiated to different programming languages. OCPL was able to leverage
-existing Iris specifications for the high-level language; for our capability
-machine model, however, it would be necessary to devise new kinds of
-specifications for our low-level programs with unstructured control-flow. It is
-likely that we could get inspiration from earlier work on logics for assembly
-programming languages, such as XCAP~\cite{xcap}.
+% In the future we would also like to
+% investigate a new program logic for reasoning about capability safety for our
+% capability machine model. We think Iris would also be a natural starting point
+% for such an endeavour, since Iris is really a framework, which can be
+% instantiated to different programming languages. OCPL was able to leverage
+% existing Iris specifications for the high-level language; for our capability
+% machine model, however, it would be necessary to devise new kinds of
+% specifications for our low-level programs with unstructured control-flow. It is
+% likely that we could get inspiration from earlier work on logics for assembly
+% programming languages, such as XCAP~\cite{xcap}.
 

### 2017-10-10 esop-2018/paper.tex
typo and restore world figure to old version

@@
 \subsection{Worlds}
-\begin{wrapfigure}{R}{0.5\textwidth}
+\begin{figure}
   \centering
-  \resizebox{6cm}{!}{\includegraphics{w11}}
+  \includegraphics{w11}
 
@@
   \label{fig:cap-world}
-\end{wrapfigure}
+\end{figure}
 % \dominique{Perhaps we could first try to convey the intuition of a permanent
@@
 in~\cite{technical_appendix}; it makes use of the isomorphism from
-Theorem~\ref{thm: world-existence}.
+Theorem~\ref{thm:world-existence}.
 

### 2017-10-11 esop-2018/paper.tex
Mention the untypedness of the logical relation in the contributions.

@@
   support: an efficient instruction for clearing a range of memory.
-\item We present a novel Kripke logical relation for reasoning about programs on the
-  capability machine. For our logical relation we prove an analogue of
-  the standard fundamental theorem of logical relations --- to the
-  best of our knowledge, our theorem
-  is % by far
-  the most general and powerful formulation of the formal guarantees offered by a
-  capability machine (a form of capability
+\item We present a novel step-indexed Kripke logical relation for reasoning
+  about programs on the capability machine. It is an untyped logical relation,
+  inspired by previous work on object capabilities \cite{Devriese:2016ObjCap}.
+  We prove an analogue of the standard fundamental theorem of logical relations
+  --- to the best of our knowledge, our theorem is % by far
+  the most general and powerful formulation of the formal guarantees offered by
+  a capability machine (a form of capability
   safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
@@
   our calling convention or a specific way of using the system's capabilities.
-  We are the first to apply these techniques for reasoning about capability machines and we believe they will prove useful for many other purposes than our calling convention.
+  We are the first to apply these techniques for reasoning about capability
+  machines and we believe they will prove useful for many other purposes than
+  our calling convention.
 \item We introduce two novel technical ideas in the unary, step-indexed Kripke

### 2017-10-12 esop-2018/paper.tex
minor changes and comments added

@@
 
-\newcommand{\sectionname}{Sec.}
+\newcommand{\sectionname}{Section}
+\renewcommand{\figurename}{Figure}
 \newcommand{\forcenewline}{$\phantom{v}$\\}
@@
 \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
-\institute{Aarhus University\email{\{lau,birkedal\}@cs.au.dk}
-\and imec-DistriNet, KU Leuven\email{dominique.devriese@cs.kuleuven.be}}
+\institute{Aarhus University - \email{\{lau,birkedal\}@cs.au.dk}
+\and imec-DistriNet, KU Leuven - \email{dominique.devriese@cs.kuleuven.be}}
 
@@
 \item We formalize a simple but representative capability machine featuring local
-  capabilities, and its operational semantics
+  capabilities and its operational semantics
   (\sectionname~\ref{sec:capab-mach-with}).
@@
 words. The state of the entire machine is represented as a configuration that is
-either a running $\Phi \in \ExecConfs$ containing a memory and a register file,
+either a running state $\Phi \in \ExecConfs$ containing a memory and a register file,
 or a failed or halted state, where the latter keeps hold of the final state of
@@
 instructions take care that the resulting capability always carries less
-authority than the original (e.g. restrict will only weaken a permission).
+authority than the original (e.g. $\restricttwo{}{}$ will only weaken a permission).
 Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a
@@
 jumped to, the instructions can start executing and have access to the local
-data through the updated version of $c$ that is then in $\pcreg$\footnote{Note
-that the capability in the $\pcreg$-register can be used to load from.}.  %
-\lau{As we % talked about what is described here, is not how we make closures in
-the TR, so % we should consider whether we should refrain from using that word
+data through the updated version of $c$ that is then in $\pcreg$. 
+\lau{As we talked about what is described here, is not how we make closures in
+the TR, so we should consider whether we should refrain from using that word
 here. }
@@
 work, we construct the activation record as depicted in
-Figure~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
+\figurename~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
 over the entire activation record (containing the previous return and stack
@@
 and a global callback for our callback to invoke. If our local state contains no
-local capabilities, then, following our calling convention, the callback would
+local capabilities, then, otherwise following our calling convention, the callback would
 not fail and the adversary could use a stored capability for the global data
@@
 %  \end{withmathindent}
-  \caption{Logical relation}
+  \caption{The logical relation.}
   \label{fig:logrel}            
@@
 kind of permission a capability can have.
-{
+
 For capabilities with read permission, the $\readCond{}$ ensures that it can
@@
 can be updated to point anywhere in their range, so they must be safe as a
-program counter (in the $\stder$-relation) no matter where they point to. In
+program counter (in the $\stder$-relation) no matter the current address. In
 contrast, enter capabilities are opaque and can only be used to jump to the
@@
 them to be safe as a program counter after the permission is changed to $\exec$.
+% This is the only condition that depend on the current address, a, because this is the 
+% only kind of capability in which the current address cannot be changed.
 Because the capabilities are not necessarily invoked immediately, this must be
@@
 
+% The FTLR does not mention integers and non-executable capabilities because they are trivially safe (it is safe to fail)
+
 \section{Examples}
@@
 The two examples also illustrate the versatility of the logical relation.
-Because it is not specific to any calling convention, we can use it to reason
+The logical relation is not specific to any calling convention, so we can use it to reason
 about both programs, even though they use different calling conventions.
@@
 \texttt{\footnotesize{f4}} returns.
-By utilizing the reinstation mechanism of the future world relation as well as our knowledge of the future worlds used, we can construct a world which the adversary's invariants are preserved.
+By utilizing the reinstation mechanism of the future world relation as well as our knowledge of the future worlds used, we can construct a world in which the adversary's invariants are preserved.
 The details of this and the proofs of the other lemmas are found in the technical appendix\cite{technical_appendix}.
@@
 \begin{comment}
-In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
+In \figurename~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
 the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
@@
 \emph{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in
-Figure~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an
+\figurename~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an
 existing pattern called biorthogonality. Biorthogonality is a pattern for

### 2017-10-17 fully-abstract-comp/technical-report.tex
Typo.
added text about decode/encode functions

@@
 \newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
-\newcommand{\scall}[2]{\twoinstr{\src{call}}{#1}{#2}}
 \newcommand{\sxjmp}[2]{\twoinstr{\src{xjmp}}{#1}{#2}}
 \newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
+% scall - special two instruction
+\newcommand{\scall}[3][]{  
+\ifthenelse{\equal{#2#3}{}}
+  {\ensuremath{\zinstr{\src{call}}_{#1}}}
+  {\ensuremath{\zinstr{\src{call}}_{#1} \; #2 \; #3}}
+}
 
@@
 \newcommand{\Word}{\shareddom{Word}}
+\newcommand{\Instr}{\shareddom{Instr}}
 \newcommand{\Mem}{\shareddom{Memory}}
@@
 \newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
+\newcommand{\exec}[1]{\plainfun{executable}{#1}}
 %\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
@@
 \newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
+\newcommand{\callCond}[1]{\plainfun{callCondition}{#1}}
 
@@
 \]
+$i$ defines the set $\Instr$
 
@@
 \]
+There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
+
+The source machine allows for variable length instructions which we utilize for $\scall{}{}$.
+In other words, when we $\scall{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction.
+This is described in more detail when we present the decoding function.
+
 \lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine.
@@
 \lau{28-09-2017: I can see from the commit history that this was something when we had push/pop instructions as well as a return instruction. I have removed the restriction on $\rstk$.}
-There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
 
@@
 
+\subsubsection{Step relations}
+\paragraph{Decode and encode functions}
+We assume functions $\decInstr{} : \Word \fun \Instr$ and $\encInstr{} : \Instr \fun \ints$ where $\Instr$ is the set of target level instructions.
+
+The $\decInstr{}$ function must be surjective and injective for all non-$\tfail$ instructions.
+The $\encInstr{}$ function must be injective.
+Futher $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
+\[
+  \decInstr{}(\encInstr{i}) = i
+\]
+
+These functions are used for both the target and source level machine. When we write instructions in places where words are required, we will assume that $\encInstr{}$ is implicit.
+
+\paragraph{Step relation}
+The instruction $\scall{}{}$ has length $\calllen$ which means that it does not fit on one memory address.
+In fact, $\scall{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
+We therefore introduce $\scall[1]{r_1}{r_2},\dots,\scall[\calllen-1]{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
+We define the following condition that indicates that $\aaddr$ is the first address of a $\scall{r_1}{r_2}$ instruction in the configuration $\Phi$:
+\[
+  \sourcecolor\left.
+    \callCond{\Phi,r_1,r_2,\aaddr} = \left\{
+      \begin{array}{l}
+        \Phi.\mem(\aaddr) = \scall[1]{r_1}{r_2} \tand\\
+        \vdots \\
+        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen]{r_1}{r_2}
+      \end{array}
+      \right.
+  \right.
+\]
+
+We use the following step relation:
+
+\begin{align*}
+  \src{\Phi} & \; \src{\step \sem{\scall{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
+                                                  \begin{array}[t]{l}
+                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
+                                                    \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
+                                                    \baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr \tand \\
+                                                    \exec{\Phi(\pcreg)}
+                                                  \end{array}\right.\\
+  \Phi & \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
+                                                  \begin{array}[t]{l}
+                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
+                                                    \neg\callCond{\Phi,r_1,r_2,\aaddr} \tand\\
+                                                    \baddr \leq \aaddr \leq \eaddr\tand\\
+                                                    \exec{\Phi(\pcreg)}
+                                                  \end{array}\right.\\
+  \Phi& \step \failed & & \totherwise
+\end{align*}
+
+\paragraph{Call implementation}
+\label{par:call-impl}
+\lau{16-10-2017: Fill this paragraph with $\scall{}{}$ implementation, i.e., the instructions $\scall[1]{r_1}{r_2} \dots \scall[\calllen]{r_1}{r_2}$ is equal to.
+Add convenience notation like $\update{\mem.a}{\scall{r_1}{r_2}} = \dots$}
+\lau{16-10-2017: Implementation of $\scall{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
+
 \subsubsection{Helpful functions and sets}
@@
   \end{cases}
+\]
+
+\[
+  \exec{\cb} \defeq \dots
 \]
@@
 \section{Compiler}
+\label{sec:compiler}
 \[
@@
   \label{def:points-to-instr}
-  A configuration $(\ms,\reg)$ {\sourcecolor or $(\ms,\reg,\_,\_)$} os said to point instruction $i$ in $A$ iff
+  A configuration $(\ms,\reg)$ {\sourcecolor or $(\ms,\reg,\_,\_)$} is said to point instruction $i$ in $A$ iff
   \[

### 2017-10-19 fully-abstract-comp/technical-report.tex
Added some VERY preliminary LR definitions along with comments here and there. The value relation is still missing entirely and some of the other definitions do not 'fit' together.

@@
 
+\newcommand{\npair}[2][n]{\left(#1,#2\right)}
+
 % Function arrows
@@
 \newcommand{\tconeq}{\mathrel{\approx_{\mathrm{ctx}}}}
+
+%%% Logical Relation notation
+\newcommand{\typesetlr}[1]{\mathcal{#1}}
+\newcommand{\lre}{\typesetlr{E}}
+\newcommand{\lrk}{\typesetlr{K}}
+\newcommand{\lrr}{\typesetlr{R}}
+\newcommand{\lro}{\typesetlr{O}}
+\newcommand{\lrv}{\typesetlr{V}}
+
+\newcommand{\stpair}[3][]{
+\ifthenelse{\equal{#1}{}}
+{\left(\src{#2}_S,#3_T\right)}
+{\left(\src{#2},#3\right)}}
+
+\newcommand{\memSat}[4][n]{#2 :_{#1}^{#3}#4}
+
+\newcommand{\World}{\mathrm{World}}
+
+\newcommand{\future}{\sqsupseteq}
+\newcommand{\privft}{\future^{\mathrm{priv}}}
+\newcommand{\pubft}{\future^{\mathrm{pub}}}
 
@@
 \newcommand{\diverge}{\Uparrow}
+\newcommand{\term}[1][-]{\Downarrow}
 
@@
                                                         &\updReg{\rstk}{c_\stk}\\
-                                                        &\updReg{r_{\var{ret,c}}}{\retptrc(\sigma_?)}\\
-                                                        &\updReg{r_{\var{ret,d}}}{\retptrd(\sigma_?)}
+                                                        &\updReg{\rretc}{\retptrc(\sigma_?)}\\
+                                                        &\updReg{\rretd}{\retptrd(\sigma_?)}
                                                  \end{array}
@@
 \clearpage
+
+\section{Logical Relation}
+\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
+\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem staisfaction).
+That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
+\[
+  \lre(W) = \left\{ \npair{\stpair{v}{v}} \middle| 
+    \begin{array}{l}
+      \forall n' \leq n \ldotp \\
+      \quad \forall \npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W) \ldotp \\
+      \qquad \forall \stpair{\reg}{\reg}\ldotp \npair[n']{
+      \left(\array{l}\src{\reg_S\updReg{\pcreg,\rretd,\rretc}{v_S,k_{d,S},k_{c,S}}},\\\reg_T\updReg{\pcreg,\rretd,\rretc}{v_T,k_{d,T},k_{c,T}}\endarray\right)} \in \lrr(W) \Rightarrow\\
+      \qquad \quad \forall \src{\ms_S},\ms_T \ldotp \memSat[n']{\stpair[.]{\ms}{\ms}}{\stpair{\reg}{\reg}}{W} \Rightarrow \\
+      \qquad \qquad \npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg}{v_S}\updReg{r_\ret}{k_V},\ms_S)},\\(\reg_T\updReg{\pcreg}{v_T}\updReg{r_\ret}{k_T},\ms_T)\endarray\right)}
+      \in \lro(W)
+    \end{array}
+            \right\}
+\]
+
+\[
+  \lro(W) = \left\{ \npair{\stpair[.]{(\reg_S,\ms_S,\stk_S,\ms_{\stk,S})}{\reg_R,\ms_R}} \middle|
+    \begin{array}{l}
+      \forall i \leq n, \stpair{\mem}{\mem} \ldotp \\
+      \quad \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\reg_R,\ms_R) \term\\
+      \quad tand\\
+      \quad (\reg_R,\ms_R) \term[i] \Rightarrow \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term
+    \end{array}
+\right\}
+\]
+
+\lau{19-10-2017: $\lrk$ may indicate that $\lre$ should really be a relation between a \src{source pc} and a pair of a target pc and stack pointer?
+Otherwise, what happens with the data part of the return pointer?}
+\[
+  \lrk(W) = \left\{ \npair{\stpair[.]{\retptrd{\sigma},\retptrc{\sigma}}{k_{d,T},k_{c,T}}} \middle|
+    \begin{array}{l}
+      k_{d,T} \text{ is a stack pointer with correct end and such with everything related to the source stack} \\
+      k_{c,T} \text{ relates to the old pc on the stack}\\
+      \forall n' \leq n \ldotp \\
+      \quad \forall W' \privft W \ldotp \\
+      \qquad \dots \in \lre(W')
+    \end{array}
+            \right\}
+\]
+
+\[
+  \lrr(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+    \begin{array}{l}
+      \exists S : \RegName \fun \World \ldotp \\
+      \quad S \text{ is a separation of $W$} \tand
+      \quad \forall r \in \dom(\reg_T) \setminus \{\pcreg, \rretd,\rretc \}\ldotp
+      \npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
+    \end{array}
+            \right\}
+\]
+Where $S\text{ is a separation of $W$}$ means that no two partion share the same ``pure'' region (whatever that is).
+
+\[
+  \lrv(W) = \left\{ \dots \right\}
+\]
+
 \section{Full Abstraction}

### 2017-10-23 fully-abstract-comp/technical-report.tex
reformatting

@@
 That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
+In the following definitions, \src{blue} is used to indicate values related to the source machine.
+This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
 \[
@@
     \begin{array}{l}
-      \forall n' \leq n \ldotp \\
-      \quad \forall \npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W) \ldotp \\
-      \qquad \forall \stpair{\reg}{\reg}\ldotp \npair[n']{
-      \left(\array{l}\src{\reg_S\updReg{\pcreg,\rretd,\rretc}{v_S,k_{d,S},k_{c,S}}},\\\reg_T\updReg{\pcreg,\rretd,\rretc}{v_T,k_{d,T},k_{c,T}}\endarray\right)} \in \lrr(W) \Rightarrow\\
-      \qquad \quad \forall \src{\ms_S},\ms_T \ldotp \memSat[n']{\stpair[.]{\ms}{\ms}}{\stpair{\reg}{\reg}}{W} \Rightarrow \\
-      \qquad \qquad \npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg}{v_S}\updReg{r_\ret}{k_V},\ms_S)},\\(\reg_T\updReg{\pcreg}{v_T}\updReg{r_\ret}{k_T},\ms_T)\endarray\right)}
+      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad\src{\reg_S}(\rretd,\rretc,\pcreg) = \src{k_{d,S}}, \src{k_{c,S}}, \src{v_S} \wedge\\
+      \quad\reg_T(\rretd,\rretc,\pcreg) = k_{d,S}, k_{c,S}, v_T \wedge\\
+      \quad\npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W) \wedge \\
+      \quad\stpair{\reg}{\reg}\ldotp \npair[n']{
+      \quad\left(\array{l}\src{\reg_S\updReg{\pcreg,\rretd,\rretc}{v_S,k_{d,S},k_{c,S}}},\\\reg_T\updReg{\pcreg,\rretd,\rretc}{v_T,k_{d,T},k_{c,T}}\endarray\right)} \in \lrr(W) \wedge\\
+      \quad\src{\ms_S},\ms_T \ldotp \memSat[n']{\stpair[.]{\ms}{\ms}}{\stpair{\reg}{\reg}}{W} \\
+      \quad\Rightarrow\\
+      \quad\npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg}{v_S}\updReg{r_\ret}{k_V},\ms_S)},\\(\reg_T\updReg{\pcreg}{v_T}\updReg{r_\ret}{k_T},\ms_T)\endarray\right)}
       \in \lro(W)
@@
       \quad \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\reg_R,\ms_R) \term\\
-      \quad tand\\
+      \quad \tand\\
       \quad (\reg_R,\ms_R) \term[i] \Rightarrow \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term
@@
       \exists S : \RegName \fun \World \ldotp \\
-      \quad S \text{ is a separation of $W$} \tand
-      \quad \forall r \in \dom(\reg_T) \setminus \{\pcreg, \rretd,\rretc \}\ldotp
-      \npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
+      \quad S \text{ is a separation of $W$} \tand\\
+      \quad \forall r \in \dom(\reg_T) \setminus \{\pcreg, \rretd,\rretc \}\ldotp\\
+      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
     \end{array}

### 2017-10-23 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
Start work on a PriSC extended abstract for the linear caps-well-bracketedness work.

\documentclass[sigplan, review]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{cleveref}
\usepackage{natbib}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
% \acmDOI{10.475/123_4}

% ISBN
% \acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[PriSC]{Workshop on Principles of Secure Compilation}{January 2018}{Los Angeles, California USA} 
\acmYear{2018}
\copyrightyear{2018}

% \acmPrice{15.00}

%\acmBadgeL[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
%\acmBadgeR[http://ctuning.org/ae/ppopp2016.html]{ae-logo}

\begin{document}
\title{Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
% \titlenote{Produces the permission block, and
%   copyright information}
\subtitle{Extended Abstract}
% \subtitlenote{The full version of the author's guide is available as
%   \texttt{acmart.pdf} document}

\author{Lau Skorstengaard}
% \orcid{}
\affiliation{%
  \institution{Aarhus University}
}
\email{lau@cs.au.dk}

\author{Dominique Devriese}
\orcid{0000-0002-3862-6856}
\affiliation{%
  \institution{imec-DistriNet, KU Leuven}
}
\email{dominique.devriese@cs.kuleuven.be}

\author{Lars Birkedal}
\affiliation{%
  \institution{Aarhus University}
}
\email{birkedal@cs.au.dk}

% The default list of authors is too long for headers}
% \renewcommand{\shortauthors}{B. Trovato et al.}

% \begin{abstract}
% This paper provides a sample of a \LaTeX\ document which conforms,
% somewhat loosely, to the formatting guidelines for
% ACM SIG Proceedings. 
% \end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
% \begin{CCSXML}
% <ccs2012>
%  <concept>
%   <concept_id>10010520.10010553.10010562</concept_id>
%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%   <concept_significance>500</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010575.10010755</concept_id>
%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
%   <concept_significance>300</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10010520.10010553.10010554</concept_id>
%   <concept_desc>Computer systems organization~Robotics</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
%  <concept>
%   <concept_id>10003033.10003083.10003095</concept_id>
%   <concept_desc>Networks~Network reliability</concept_desc>
%   <concept_significance>100</concept_significance>
%  </concept>
% </ccs2012>  
% \end{CCSXML}

% \ccsdesc[500]{Computer systems organization~Embedded systems}
% \ccsdesc[300]{Computer systems organization~Redundancy}
% \ccsdesc{Computer systems organization~Robotics}
% \ccsdesc[100]{Networks~Network reliability}

% \keywords{ACM proceedings, \LaTeX, text tagging}

% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{sampleteaser}
%   \caption{This is a teaser}
%   \label{fig:teaser}
% \end{teaserfigure}

\maketitle

\section{Introduction}
Secure compilers preserve source-language (security-relevant) properties even when the  compiled code interacts with arbitrary target-language components.
Generally, properties that hold in the source language but not in the target language, need to be somehow enforced by the compiler.
Two properties that hold in many high-level source languages, but not in the assembly languages that they are compiled to are the well-bracketedness of control flow and the encapsulation of local state.

Well-bracketed control flow expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations. 
At the assembly level, this is not so obvious, as invoked functions get direct access to return pointers, that they are supposed to jump to a single time, at the end of their execution, but there is no guarantee that untrusted assembly code respects this intended usage.
Particularly, it may invoke return pointers from other stack frames than the one it was invoked from, either ones from frames higher in the call stack or from stack frames that no longer exist because they have already been returned to before.

Local state encapsulation is the guarantee that when a function invokes another function, its local variables will not have been modified when the invoked function returns.
At the assembly level, this property is also far from obvious, local variables are stored on the stack when another function is invoked, and functions are not supposed to touch stack frames other than their own.
However, untrusted assembly code is usually not prevented from ignoring this requirement and overwriting the local state of its caller or other stack frames.

To enforce these properties, a target language is needed with some form of security primitive that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.

\section{Conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}

### 2017-10-23 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
Some more work on the PriSC extended abstract.

@@
 To enforce these properties, a target language is needed with some form of security primitive that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
+The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this type of enforcement.
+More suitable security primitives are offered by a type of processors known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
+These processors use tagged memory to enforce a strict separation between integers and \emph{capabilities}: pointers that carry authority.
+Capabilities come in different flavours.
+Memory capabilities are used to read and write to memory.
+Additionally, CHERI offers a notion of sealed code-data pairs of capabilities that represent a type of encapsulated closure: a piece of code coupled with private state that it gains access to upon invocation.\footnote{Other capability machines offer other mechanisms that can be used to the same effect.}
+
+CheriBSD, an operating sytem built on the CHERI capability machine, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
+To prevent the stack pointer (or derivatives from it) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
+The literature does not contain many details about the mechanism, let alone a formal analysis, but most importantly, the mechanism does not seem like it can support higher-order interfaces (e.g.
+C function pointers) in a scalable fashion.
+
+In recent work, we have been looking at an alternative approach \citep{skorstengaard_reasoning_nodate}.
+It is also based on CHERI's local capabilities, but combines them with a more traditional, single, shared stack that scales better in the presence of higher-order interfaces.
+By passing stack and return pointers as local capabilities, our mechanism effectively prevents callees from storing them on the heap.
+Using a step-indexed Kripke logical relation, we are able to prove correctness of programs that rely on well-bracketed control flow in a complex way.
+
+While this approach works quite well, this paper investigates another alternative approach.
+The aim of this new approach is to resolve two limitations: (1) the need to clear the stack on boundary crossings and (2) the lack of a formal statement of well-bracketed control flow and local state encapsulation.
+In the next two sections, we discuss these two limitations and how we intend to resolve them.
+
+\section{Linear Capabilities, not Local}
+
+\section{Formalising well-bracketed control flow}
 
 \section{Conclusion}
+Contributions here?
 

### 2017-10-24 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
Finish first draft of PriSC submission.

@@
 \usepackage{natbib}
+\usepackage[utf8]{inputenc}
 
@@
 % \subtitlenote{The full version of the author's guide is available as
-%   \texttt{acmart.pdf} document}
+%   \texttt{acDat zou in dmart.pdf} document}
 
@@
 
-In recent work, we have been looking at an alternative approach \citep{skorstengaard_reasoning_nodate}.
+In recent work, we have been looking at an alternative approach \citep{skorstengaard_reasoning_2017}.
 It is also based on CHERI's local capabilities, but combines them with a more traditional, single, shared stack that scales better in the presence of higher-order interfaces.
@@
 \section{Linear Capabilities, not Local}
+What we essentially want to achieve in a secure calling convention is to give an invoked function access to a stack pointer and a return pointer, but only temporarily: during the execution of the invoked function.
+In our original approach, we make both pointers local, so that the CHERI processor guarantees, essentially, that they can only be stored in registers or on the stack.
+To make sure that an invoked function has no way of transferring the pointers accross outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
+In other words, our approach relies on clearing registers and the stack before we pass control to untrusted code, as a way of preventing it from obtaining access to other stack and return pointers than the ones for the stack frame it is invoked in.
+The requirement for clearing registers is unproblematic, but clearing the entire stack upon every boundary crossing is a non-trivial requirement.
+We speculate that it can be made efficient with custom hardware extensions\footnote{Specifically, we imagine a kind of pre-L0 processor cache that does not retain values but only remembers that a certain range has been zeroed out.}, but this is not obvious.
+
+An alternative that does not have this requirement, relies on \emph{linear capabilities}, rather than local ones.
+Linear capabilities are (to the best of our knowledge) a new concept, similar to CHERI's local capabilities.\footnote{One author of this abstract is also investigating linear capabilities in another application, also submitted to PriSC.}
+The idea is that all capabilities are marked as either regular or linear.
+When they are linear, they are subject to special restrictions, enforced by the hardware, that prevent them from being duplicated.
+When linear capabilities are moved to a new location in registers or memory, they are erased from their old location.
+A few special instructions are needed to make this workable, e.g.\ a split instruction that splits a memory capability for a region of memory into disjoint capabilities for two parts of the region and a splice instruction that performs the reverse.
+Linearity of a capability is a powerful tool: when we pass a linear capability to untrusted code and we receive the same capability back, we can be sure that they will not have been able to store a copy of it.
+Similarly, if we receive a linear capability from untrusted code and we know that the capability addresses memory that is only ever addressed by linear capabilities, then we are sure that the untrusted code can no longer access that memory.
+
+As such, linear capabilities can be used as an alternative to local capabilities to enforce well-bracketed control flow.
+We cannot provide full details here for space reasons, but the idea is essentially the following.
+By passing stack and return pointers as linear capabilities, and requiring that invoked functions hand in their stack pointer upon return, untrusted code are prevented from keeping hold of old stack and return pointers past their intended lifetime, essentially because they can only ever have one copy that they have to hand in when this lifetime ends.
 
 \section{Formalising well-bracketed control flow}
+The second improvement we are making is related to the properties of well-bracketed control flow and local state encapsulation.
+Our previous work did not actually have a precise statement of these properties.
+We only show a reasoning technique that can be used to prove certain example programs that rely on the properties.
+Although it can be argued that this proof technique should perhaps qualify as a formal statement of the properties, we are investigating an alternative approach, based on the notion of fully abstract compilation \citep{abadi_protection_1999}.
 
-\section{Conclusion}
-Contributions here?
+We intend to start from a formalisation of an assembly language for a capability machine with linear capabilities (the target language).
+Our idea is then to introduce a variant of this target language that features a primitive stack, built into the operational semantics, as well as additional instructions like call and return.
+While the original target language is a standard assembly language without even the notion of a stack, this variant language (the source language) does have a stack, and control flow well-bracketedness and local state encapsulation are evident properties of the operational semantics.
+We can formalise our calling convention in the definition of a compiler from the source language to the target language.
+Additionally, we conjecture that this compiler can be proved fully abstract, thus formally expressing and establishing security of the calling convention.
+
+Interestingly, the proof of full abstraction for this compiler is expected to use a very simple back-translation, that simply maps every target language instruction to the corresponding instruction in the source.
+This back-translation works because we construct our source language in such a way that it has direct counterparts to every target language value.
+Specifically, the source language contains a representation of stack and return pointers as unspecified abstract values whose behavior corresponds to their target counterparts, but in a way that guarantees the intended properties of well-bracketed control flow and local state encapsulation.
 

### 2017-10-24 fully-abstract-comp/technical-report.tex
LR updates

@@
 \newcommand{\lrv}{\typesetlr{V}}
+\newcommand{\lrp}{\typesetlr{P}}
+\newcommand{\lrm}{\typesetlr{M}}
 
@@
 \ifthenelse{\equal{#1}{}}
-{\left(\src{#2}_S,#3_T\right)}
+{\left(\src{#2_S},#3_T\right)}
 {\left(\src{#2},#3\right)}}
 
-\newcommand{\memSat}[4][n]{#2 :_{#1}^{#3}#4}
+\newcommand{\memSat}[3][n]{#2 :_{#1}#3}
 
@@
       \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad\src{\reg_S}(\rretd,\rretc,\pcreg) = \src{k_{d,S}}, \src{k_{c,S}}, \src{v_S} \wedge\\
-      \quad\reg_T(\rretd,\rretc,\pcreg) = k_{d,S}, k_{c,S}, v_T \wedge\\
-      \quad\npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W) \wedge \\
-      \quad\stpair{\reg}{\reg}\ldotp \npair[n']{
-      \quad\left(\array{l}\src{\reg_S\updReg{\pcreg,\rretd,\rretc}{v_S,k_{d,S},k_{c,S}}},\\\reg_T\updReg{\pcreg,\rretd,\rretc}{v_T,k_{d,T},k_{c,T}}\endarray\right)} \in \lrr(W) \wedge\\
-      \quad\src{\ms_S},\ms_T \ldotp \memSat[n']{\stpair[.]{\ms}{\ms}}{\stpair{\reg}{\reg}}{W} \\
-      \quad\Rightarrow\\
-      \quad\npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg}{v_S}\updReg{r_\ret}{k_V},\ms_S)},\\(\reg_T\updReg{\pcreg}{v_T}\updReg{r_\ret}{k_T},\ms_T)\endarray\right)}
+      \quad \exists W_\lrp, W_\lrk, W_\lrr, W_\lrm \text{that is a separarion of $W$}\ldotp  \wedge \\
+      \qquad\npair[n']{\stpair{v}{v}} \in \lrp(W_\lrp) \wedge \\
+      \qquad\npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W_\lrk) \wedge \\
+      \qquad\npair[n']{\stpair[]{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\Rightarrow\\
+      \qquad\npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg,\rretc,\rretd}{v_S,k_{c,S},k_{d,S}},\ms_S,\stk, \ms_\stk)},\\
+                                          (\reg_T\updReg{\pcreg,\rretc,\rretd}{v_T,k_{c,T},k_{d,T}},\ms_T)\endarray\right)}
       \in \lro(W)
@@
 \]
+\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out.
+I can see how it will enforce linearity, but is you give a "pure" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part.
+In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file.
+Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}
 
@@
     \end{array}
+            \right\}
+\]
+
+\[
+  \lrp(W) = \left\{
+    \npair{\stpair{v}{v}} \middle| \dots
   \right\}

### 2017-10-24 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
typos

@@
 
-CheriBSD, an operating sytem built on the CHERI capability machine, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
+CheriBSD, an operating system built on the CHERI capability machine, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
 To prevent the stack pointer (or derivatives from it) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
@@
 In our original approach, we make both pointers local, so that the CHERI processor guarantees, essentially, that they can only be stored in registers or on the stack.
-To make sure that an invoked function has no way of transferring the pointers accross outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
+To make sure that an invoked function has no way of transferring the pointers across outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
 In other words, our approach relies on clearing registers and the stack before we pass control to untrusted code, as a way of preventing it from obtaining access to other stack and return pointers than the ones for the stack frame it is invoked in.

### 2017-10-25 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
Address remarks by Lau and Lars in PriSC extended abstract.

@@
 The second improvement we are making is related to the properties of well-bracketed control flow and local state encapsulation.
-Our previous work did not actually have a precise statement of these properties.
-We only show a reasoning technique that can be used to prove certain example programs that rely on the properties.
-Although it can be argued that this proof technique should perhaps qualify as a formal statement of the properties, we are investigating an alternative approach, based on the notion of fully abstract compilation \citep{abadi_protection_1999}.
+To establish these properties, our previous work followed the approach taken by previous work in other settings \citep{dreyer_impact_2010}: we provide a sound reasoning technique and use it to prove soundness of challenging example programs that rely on the properties.
+In the current work, we are investigating an alternative way to formalise the two properties, based on the notion of fully abstract compilation \citep{abadi_protection_1999}.
 
 We intend to start from a formalisation of an assembly language for a capability machine with linear capabilities (the target language).
-Our idea is then to introduce a variant of this target language that features a primitive stack, built into the operational semantics, as well as additional instructions like call and return.
-While the original target language is a standard assembly language without even the notion of a stack, this variant language (the source language) does have a stack, and control flow well-bracketedness and local state encapsulation are evident properties of the operational semantics.
+Our idea is then to introduce a variant of this target language that features a primitive stack, built into the operational semantics, as well as additional instructions like call.
+While the original target language is a standard assembly language without even the notion of a stack, this variant language (the source language) does have a stack, and control flow well-bracketedness and local state encapsulation follow more directly from the operational semantics.
 We can formalise our calling convention in the definition of a compiler from the source language to the target language.

### 2017-10-25 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex

@@
 \usepackage{cleveref}
+\usepackage{hyperref}
 \usepackage{natbib}
@@
 
-% \begin{abstract}
-% This paper provides a sample of a \LaTeX\ document which conforms,
-% somewhat loosely, to the formatting guidelines for
-% ACM SIG Proceedings. 
-% \end{abstract}
+\begin{abstract}
+We propose and study a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
+The calling convention is based on linear capabilities, a new type of capabilities that are prevented by the hardware from being duplicated.
+In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, building on the concept of fully abstract compilation.
+\end{abstract}
 
@@
 Generally, properties that hold in the source language but not in the target language, need to be somehow enforced by the compiler.
-Two properties that hold in many high-level source languages, but not in the assembly languages that they are compiled to are the well-bracketedness of control flow and the encapsulation of local state.
+Two properties that hold in many high-level source languages, but not in the assembly languages they are compiled to, are well-bracketed control flow and encapsulation of local state.
 
@@
 At the assembly level, this is not so obvious, as invoked functions get direct access to return pointers, that they are supposed to jump to a single time, at the end of their execution, but there is no guarantee that untrusted assembly code respects this intended usage.
-Particularly, it may invoke return pointers from other stack frames than the one it was invoked from, either ones from frames higher in the call stack or from stack frames that no longer exist because they have already been returned to before.
+Particularly, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned before.
 
 Local state encapsulation is the guarantee that when a function invokes another function, its local variables will not have been modified when the invoked function returns.
-At the assembly level, this property is also far from obvious, local variables are stored on the stack when another function is invoked, and functions are not supposed to touch stack frames other than their own.
-However, untrusted assembly code is usually not prevented from ignoring this requirement and overwriting the local state of its caller or other stack frames.
+At the assembly level, this property is also far from obvious.
+The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
+However, untrusted assembly code is free to ignore this requirement and overwrite the local state of its caller or other stack frames.
 
-To enforce these properties, a target language is needed with some form of security primitive that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
-The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this type of enforcement.
+To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
+The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
 More suitable security primitives are offered by a type of processors known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
@@
 Capabilities come in different flavours.
-Memory capabilities are used to read and write to memory.
-Additionally, CHERI offers a notion of sealed code-data pairs of capabilities that represent a type of encapsulated closure: a piece of code coupled with private state that it gains access to upon invocation.\footnote{Other capability machines offer other mechanisms that can be used to the same effect.}
+Memory capabilities are used to read from and write to memory.
+Additionally, the CHERI capability machine offers sealed code-data pairs of capabilities that represent an encapsulated closure: a piece of code coupled with private state that it gains access to upon invocation.\footnote{Other capability machines offer other mechanisms that can be used to the same effect.}
 
-CheriBSD, an operating system built on the CHERI capability machine, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
-To prevent the stack pointer (or derivatives from it) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
-The literature does not contain many details about the mechanism, let alone a formal analysis, but most importantly, the mechanism does not seem like it can support higher-order interfaces (e.g.
+CheriBSD, an operating system built on CHERI, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
+To prevent the stack pointer (or derivatives) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
+The literature does not contain many details about the mechanism, let alone a formal analysis, but additionally, the mechanism does not seem able to support higher-order interfaces (e.g.
 C function pointers) in a scalable fashion.
@@
 In recent work, we have been looking at an alternative approach \citep{skorstengaard_reasoning_2017}.
-It is also based on CHERI's local capabilities, but combines them with a more traditional, single, shared stack that scales better in the presence of higher-order interfaces.
-By passing stack and return pointers as local capabilities, our mechanism effectively prevents callees from storing them on the heap.
+We also use CHERI's local capabilities, but combine them with a more traditional, single, shared stack that scales better in the presence of higher-order interfaces.
+By passing stack and return pointers as local capabilities, we effectively prevent callees from storing them on the heap.
 Using a step-indexed Kripke logical relation, we are able to prove correctness of programs that rely on well-bracketed control flow in a complex way.
@@
 While this approach works quite well, this paper investigates another alternative approach.
-The aim of this new approach is to resolve two limitations: (1) the need to clear the stack on boundary crossings and (2) the lack of a formal statement of well-bracketed control flow and local state encapsulation.
-In the next two sections, we discuss these two limitations and how we intend to resolve them.
+We aim to make two improvements, discussed in the next sections: (1) remove the need to clear the stack on boundary crossings and (2) explore a new formal statement of well-bracketed control flow and local state encapsulation.
 
@@
 In our original approach, we make both pointers local, so that the CHERI processor guarantees, essentially, that they can only be stored in registers or on the stack.
-To make sure that an invoked function has no way of transferring the pointers across outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
-In other words, our approach relies on clearing registers and the stack before we pass control to untrusted code, as a way of preventing it from obtaining access to other stack and return pointers than the ones for the stack frame it is invoked in.
+To prevent an invoked function from transferring the pointers across outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
+In other words, we need to clear registers and the stack before passing control to untrusted code, to prevent it from obtaining access to stack and return pointers for other stack frames than its own.
 The requirement for clearing registers is unproblematic, but clearing the entire stack upon every boundary crossing is a non-trivial requirement.
@@
 The idea is that all capabilities are marked as either regular or linear.
-When they are linear, they are subject to special restrictions, enforced by the hardware, that prevent them from being duplicated.
-When linear capabilities are moved to a new location in registers or memory, they are erased from their old location.
-A few special instructions are needed to make this workable, e.g.\ a split instruction that splits a memory capability for a region of memory into disjoint capabilities for two parts of the region and a splice instruction that performs the reverse.
-Linearity of a capability is a powerful tool: when we pass a linear capability to untrusted code and we receive the same capability back, we can be sure that they will not have been able to store a copy of it.
-Similarly, if we receive a linear capability from untrusted code and we know that the capability addresses memory that is only ever addressed by linear capabilities, then we are sure that the untrusted code can no longer access that memory.
+Linear ones are subject to special hardware-enforced restrictions, that prevent them from being duplicated.
+For example, when linear capabilities are moved to a new location in registers or memory, they are erased from their old location.
+A few special instructions are needed to make this workable, e.g.\ a split instruction that splits a memory capability for a region of memory into disjoint capabilities for two parts of the region and a splice instruction that does the reverse.
+This linearity is a powerful tool: when passing a linear capability to untrusted code and receiving the same capability back, we are sure that they cannot have stored a copy of it.
+Similarly, when we receive, from untrusted code, a linear capability that addresses memory that we know is only ever addressed by linear capabilities, then we are sure that the untrusted code no longer has access that memory.
 
@@
 We cannot provide full details here for space reasons, but the idea is essentially the following.
-By passing stack and return pointers as linear capabilities, and requiring that invoked functions hand in their stack pointer upon return, untrusted code are prevented from keeping hold of old stack and return pointers past their intended lifetime, essentially because they can only ever have one copy that they have to hand in when this lifetime ends.
+By passing stack and return pointers as linear capabilities, and requiring that invoked functions hand in their stack pointer upon return, untrusted code is prevented from holding on to stack and return pointers past their intended lifetime, essentially because they can only ever have one copy and they have to hand it in when this lifetime ends.
 
@@
 The second improvement we are making is related to the properties of well-bracketed control flow and local state encapsulation.
-To establish these properties, our previous work followed the approach taken by previous work in other settings \citep{dreyer_impact_2010}: we provide a sound reasoning technique and use it to prove soundness of challenging example programs that rely on the properties.
+To establish these properties, our previous work followed previous work \citep{dreyer_impact_2010} in providing a sound reasoning technique and using it to prove soundness of challenging example programs that rely on the properties.
 In the current work, we are investigating an alternative way to formalise the two properties, based on the notion of fully abstract compilation \citep{abadi_protection_1999}.
 
-We intend to start from a formalisation of an assembly language for a capability machine with linear capabilities (the target language).
-Our idea is then to introduce a variant of this target language that features a primitive stack, built into the operational semantics, as well as additional instructions like call.
-While the original target language is a standard assembly language without even the notion of a stack, this variant language (the source language) does have a stack, and control flow well-bracketedness and local state encapsulation follow more directly from the operational semantics.
-We can formalise our calling convention in the definition of a compiler from the source language to the target language.
+We intend to start by formalising the assembly language for a capability machine with linear capabilities (the target language).
+Then, we will introduce the source language: a variant of this target language that features a primitive stack, built into the operational semantics, as well as additional instructions like call.
+While the target language is a standard assembly language without even the notion of a stack, the source language does have a stack, and control flow well-bracketedness and local state encapsulation follow more directly from the operational semantics.
+We will formalise our calling convention in the definition of a compiler from the source to the target language.
 Additionally, we conjecture that this compiler can be proved fully abstract, thus formally expressing and establishing security of the calling convention.

### 2017-10-25 fully-abstract-comp/technical-report.tex
a few additions - the LR is still a mess.
updates to step relation to take special case into consideration
A3
Propose a new magic-wand-style quantifications in the expression relation.
made several small changes based on meeting with Dominique

@@
-\documentclass[a4paper]{article}
-\usepackage[a4paper]{geometry}
+\documentclass[a3paper]{article}
+\usepackage[a3paper]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \newcommand{\steps}{\step^*}
-\newcommand{\diverge}{\Uparrow}
-\newcommand{\term}[1][-]{\Downarrow}
+\newcommand{\diverge}{{\Uparrow}}
+\newcommand{\term}[1][-]{{\Downarrow_{#1}}}
 
@@
 %\newcommand{\nopermo}{\plainperm{0o}}
-\newcommand{\enter}{\plainperm{e}}
+%\newcommand{\enter}{\plainperm{e}}
 %\newcommand{\entero}{\plainperm{eo}}
@@
 \newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
-\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
+%\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
 \newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
@@
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma) \mid \retptrc(\sigma)}\\ 
+    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma,(\permbnf,\basebnf,\aendbnf,\addrbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\
     &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
@@
     \node[main node] (rwx) {$\rwx$};
-    \node[main node] (rx) [below of=rwx] {$\rx$};
-    \node[main node] (e) [below of=rx] {$\enter$};
-
-    \node[main node] (rw) [below left of=rwx,xshift=-0.5cm,yshift=-0.3cm] {$\rw$};
-    \node[main node] (r) [below of=rw] {$\readonly$};
+    \node[main node] (rx) [below right of=rwx] {$\rx$};
+
+    \node[main node] (rw) [below left of=rwx,] {$\rw$};
+    \node[main node] (r) [below right of=rw] {$\readonly$};
     \node[main node] (0) [below of=r] {$\noperm$};
@@
     (rwx) edge (rx)
-    (rx) edge (e)
 
     (rw) edge (rwx)
-    (r) edge (rx)
-    (0) edge (e);
+    (r) edge (rx);
   \end{tikzpicture}
@@
                                                     \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
-                                                    \neg\callCond{\Phi,r_1,r_2,\aaddr} \tand\\
+                                                    \src{(\neg\callCond{\Phi,r_1,r_2,\aaddr} \tor}\\
+                                                    \src{\eaddr < \aaddr + \calllen-1) \tand} \\
                                                     \baddr \leq \aaddr \leq \eaddr\tand\\
@@
   \updPcAddr{c} \defeq \dots
-\]
-
-\[
-  \updPcPerm{c} \defeq
-  \begin{cases}
-    ((\rx,\lin),\baddr,\eaddr,\aaddr) & c = ((\enter,\lin),\baddr,\eaddr,\aaddr)\\
-    c & \totherwise
-  \end{cases}
 \]
@@
                      \begin{cases}
-                       \Phi\updReg{r}{w}\updReg{\pcreg}{\updPcPerm{\Phi(r)}} & w = \linCons{\Phi(r)}
+                       \Phi\updReg{r}{w}\updReg{\pcreg}{\Phi(r)} & w = \linCons{\Phi(r)}
                      \end{cases}
@@
                                       \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                      \perm \neq \enter \tand \\
                                       c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
@@
                                       \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                      \perm \neq \enter \tand \\
                                       c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
@@
                                       \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
-                                      \perm \neq \enter \tand \\
                                       w_2 = \linCons{\Phi(r_2)}\\
@@
                                    \Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                   \perm \neq \enter \tand \\
                                    \Phi(r_4) = n \tor r_4 = n \\
@@
                                    \Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                   \perm \neq \enter \\
                                    \Phi(r_4) = n \tor r_4 = n \\
@@
                                   \Phi(r_3) = ((\perm,\lin),\baddr_3,\eaddr_3,\aaddr_3) \tand \\
-                                  \perm \neq \enter \tand \\
                                   \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
@@
                                   \Phi(r_3) = \stkptr{\perm,\baddr_3,\eaddr_3,\aaddr_3} \tand \\
-                                  \perm \neq \enter \tand \\
                                   \eaddr_2 + 1 = \baddr_3 \tand \baddr_2 \leq \eaddr_2 \tand \baddr_3 \leq \eaddr_3 \tand \\
@@
 \[
-  \lre(W) = \left\{ \npair{\stpair{v}{v}} \middle| 
+  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
     \begin{array}{l}
       \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \exists W_\lrp, W_\lrk, W_\lrr, W_\lrm \text{that is a separarion of $W$}\ldotp  \wedge \\
-      \qquad\npair[n']{\stpair{v}{v}} \in \lrp(W_\lrp) \wedge \\
-      \qquad\npair[n']{\stpair[.]{(k_{d,S},k_{c,S})}{(k_{d,T},k_{c,T})}} \in \lrk(W_\lrk) \wedge \\
-      \qquad\npair[n']{\stpair[]{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \in \lrp(W) \wedge \\
+      \qquad\npair[n']{\stpair[.]{k_{d,S},k_{c,S}}{k_{d,T},k_{c,T}}} \in \lrk(W_\lrk) \wedge \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
       \qquad\Rightarrow\\
-      \qquad\npair[n']{\left(\array{l}\src{(\reg_S\updReg{\pcreg,\rretc,\rretd}{v_S,k_{c,S},k_{d,S}},\ms_S,\stk, \ms_\stk)},\\
-                                          (\reg_T\updReg{\pcreg,\rretc,\rretd}{v_T,k_{c,T},k_{d,T}},\ms_T)\endarray\right)}
-      \in \lro(W)
+      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
+                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
+,T},k_{c,T},k_{d,T}})\endarray\right)}
+      \in \lro(W\uplus W_\lrk \uplus W_\lrr \uplus W_\lrm)
     \end{array}
@@
 \[
-  \lro(W) = \left\{ \npair{\stpair[.]{(\reg_S,\ms_S,\stk_S,\ms_{\stk,S})}{\reg_R,\ms_R}} \middle|
+  \lro(W) = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
-      \forall i \leq n, \stpair{\mem}{\mem} \ldotp \\
-      \quad \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\reg_R,\ms_R) \term\\
+      \forall i \leq n \ldotp \\
+      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T,\reg_T) \term\\
       \quad \tand\\
-      \quad (\reg_R,\ms_R) \term[i] \Rightarrow \src{(\reg_S,\ms_S \cup \mem_S,\stk_S,\ms_{\stk,S})} \term
+      \quad (\ms_T,\reg_T) \term[i] \Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \term
     \end{array}
@@
 \[
-  \lrk(W) = \left\{ \npair{\stpair[.]{\retptrd{\sigma},\retptrc{\sigma}}{k_{d,T},k_{c,T}}} \middle|
+  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma)},\\\src{\retptrc(\sigma)},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
     \begin{array}{l}
@@
       k_{c,T} \text{ relates to the old pc on the stack}\\
-      \forall n' \leq n \ldotp \\
-      \quad \forall W' \privft W \ldotp \\
-      \qquad \dots \in \lre(W')
+      \forall n' \leq n, W' \pubft W \ldotp \\
+      \quad \forall, W_\lrr, W_\lrm \ldotp  \\
+      \qquad\npair[n']{\src{\retptrd(\sigma)},\src{\retptrc(\sigma)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\Rightarrow\\
+      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\dots}},\stk, \ms_\stk)},\\
+                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\dots,\dots})\endarray\right)}
+      \in \lro(W' \uplus W_\lrr \uplus W_\lrm)
     \end{array}
@@
   \lrp(W) = \left\{
-    \npair{\stpair{v}{v}} \middle| \dots
+    \npair{\stpair{v}{v}} \middle| \text{make sure linearity constraint satisfied}\dots 
   \right\}
@@
 \[
-  \lrv(W) = \left\{ \dots \right\}
+  \lrv(W) =
+  \begin{array}[t]{l}
+    \left\{ \npair{\stpair[.]{i}{i}} \middle| i \in \ints \right\}\cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \middle| 
+    \begin{array}{l}
+    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrd(\sigma)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W') \ldotp\\
+      \npair[n']{\retptrc(\sigma),\retptrd(\sigma),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
+    \end{array}
+    \right\}\cup\\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \middle| 
+    \begin{array}{l}
+    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrc(\sigma)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W') \ldotp\\
+      \npair[n']{\src{\retptrc(\sigma)},\src{\retptrd(\sigma)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
+    \end{array}
+    \right\}\cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \middle| \right)}
+    \right\}
+  \end{array}
 \]

### 2017-10-25 fully-abstract-comp/technical-report.tex
added retptr data some places where it was missing

@@
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma,(\permbnf,\basebnf,\aendbnf,\addrbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\
+    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
     &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
@@
       \cb = \stkptr{\_,\_,\_,\_}\\
-      \cb = \retptrd(\_)
+      \cb = \retptrd(\_,\_)
     \end{array}\right.\\
@@
                                        \begin{array}[t]{l}
-                                         \Phi(r_1) = \retptrc(\sigma) \tand \\
-                                         \Phi(r_2) = \retptrd(\sigma) \tand \\
+                                         \Phi(r_1) = \retptrc(\sigma,\_) \tand \\
+                                         \Phi(r_2) = \retptrd(\sigma,\_) \tand \\
                                          \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
@@
 \lau{19-09-2017: In which cases do we have undefined here?}
+\lau{25-10-2017: I have forgotten why we save $e_\stk$ on the stack and whether this is still necessary (it at least prevent the empty case, but I do not recall if it is necessary for other reasons).}
 This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack.
@@
                                                         &\updReg{\rstk}{c_\stk}\\
-                                                        &\updReg{\rretc}{\retptrc(\sigma_?)}\\
-                                                        &\updReg{\rretd}{\retptrd(\sigma_?)}
+                                                        &\updReg{\rretc}{\retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))}\\
+                                                        &\updReg{\rretd}{\retptrd(\sigma_?,(\aaddr_\stk,\eaddr_\stk))}
                                                  \end{array}
@@
                                                    \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
-                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
+                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
                                                    \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk - 1]} \tand \\
@@
   \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
-  \item $\src{w = \retptrc{\_}}$
-  \item $\src{w = \retptrd{\_}}$
+  \item $\src{w = \retptrc(\_,\_)}$
+  \item $\src{w = \retptrd(\_,\_)}$
   \end{itemize}
@@
 \[
-  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma)},\\\src{\retptrc(\sigma)},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
+  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,\_)},\\\src{\retptrc(\sigma,\_)},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
     \begin{array}{l}
@@
       \quad \forall, W_\lrr, W_\lrm \ldotp  \\
-      \qquad\npair[n']{\src{\retptrd(\sigma)},\src{\retptrc(\sigma)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
+      \qquad\npair[n']{\src{\retptrd(\sigma,\_)},\src{\retptrc(\sigma,\_)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
@@
     \left\{ \npair{\stpair[.]{i}{i}} \middle| i \in \ints \right\}\cup \\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \middle| 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\_)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \middle| 
     \begin{array}{l}
-    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrd(\sigma)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W') \ldotp\\
-      \npair[n']{\retptrc(\sigma),\retptrd(\sigma),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
+    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W') \ldotp\\
+      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
     \end{array}
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \middle| 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\_)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \middle| 
     \begin{array}{l}
-    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrc(\sigma)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W') \ldotp\\
-      \npair[n']{\src{\retptrc(\sigma)},\src{\retptrd(\sigma)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
+    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W') \ldotp\\
+      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
     \end{array}

### 2017-10-26 fully-abstract-comp/technical-report.tex
updated value relation
added some stuff on the worlds

@@
 \newcommand{\World}{\mathrm{World}}
+\newcommand{\RegionName}{\mathrm{RegionName}}
+\newcommand{\Region}{\mathrm{Region}}
+\newcommand{\pureReg}{\mathrm{pure}}
+\newcommand{\normalReg}{\mathrm{normal}}
+\newcommand{\State}{\mathrm{State}}
+\newcommand{\Rels}{\mathrm{Rels}}
+\newcommand{\UPred}[1]{\mathrm(#1)}
 
@@
 \section{Logical Relation}
+\subsection{Worlds}
+\[
+  \World = \RegionName \parfun 2 \times \Region
+\]
+where $\RegionName = \nats$
+\[
+  \Region = \left\{
+  \begin{array}{l}
+    \{\pureReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\normalReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+  \end{array} \right.
+\]
+\lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
+\[
+  W_1 \uplus W_1 = W
+  \begin{array}[t]{l}
+    \text{if $\dom(W) = \dom(W_1) = \dom(W_2) \tand$} \\
+    \forall r \in \dom(W) \ldotp W(r) = W_1(r) \oplus W_2(r)
+  \end{array}
+\]
+$\oplus$ is defined as follows
+\begin{align*}
+  (i,\normalReg,\var{sm}) \uplus (j,\normalReg,\var{sm}) = & (\max(i,j),\normalReg,\var{sm}) \\
+  (1,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) = & (1,\pureReg,\var{sm}) \\
+  (0,\pureReg,\var{sm}) \uplus (1,\pureReg,\var{sm}) = & (1,\pureReg,\var{sm}) \\
+  (0,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) = & (0,\pureReg,\var{sm}) \\
+\end{align*}
+and for all other cases $\oplus$ is undefined.
+Specifically, $\uplus$ is not defined when both sides are an owned $\pureReg$-region.
+It is further not defined if the two sides do not agree on region type and state machine.
+
+\subsection{Relation}
 \lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
@@
     \begin{array}{l}
-    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W') \ldotp\\
-      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
+    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W_\lrv) \ldotp\\
+      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \uplus W_\lrv)
     \end{array}
@@
     \begin{array}{l}
-    \forall n'<n, W' \pubft W, \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W') \ldotp\\
-      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W')
+    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W_\lrv) \ldotp\\
+      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \uplus W_\lrv)
     \end{array}
     \right\}\cup \\
+    \left\{ \npair{\left(\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \middle|
+    \begin{array}{l}
+
+    \end{array}
+    \right\}\cup\\
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \middle| \right)}

### 2017-10-27 fully-abstract-comp/technical-report.tex
A bit more text about the proposed worlds

@@
 \]
-where $\RegionName = \nats$
+where $\RegionName = \nats$. $2$ indicates ownership of the region which decides whether a capability is allowed to reference this part of memory.
+\lau{26-10-2017: Maybe we want to place the ownership directly on the $\pureReg$-regions as we only want to enforce that linear capabilities don't have any aliases.}
 \[
@@
     \{\pureReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
-    \{\normalReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\normalReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
   \end{array} \right.
 \]
+\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitely into the worlds (but maybe we can handle it in the memory satisfaction).}
 \lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
+
+Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\pureReg$-regions.
 \[
@@
 \begin{align*}
-  (i,\normalReg,\var{sm}) \uplus (j,\normalReg,\var{sm}) = & (\max(i,j),\normalReg,\var{sm}) \\
-  (1,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) = & (1,\pureReg,\var{sm}) \\
-  (0,\pureReg,\var{sm}) \uplus (1,\pureReg,\var{sm}) = & (1,\pureReg,\var{sm}) \\
-  (0,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) = & (0,\pureReg,\var{sm}) \\
+  (i,\normalReg,\var{sm}) \uplus (j,\normalReg,\var{sm}) =  & \; (\max(i,j),\normalReg,\var{sm}) \\
+  (1,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) =      & \; (1,\pureReg,\var{sm}) \\
+  (0,\pureReg,\var{sm}) \uplus (1,\pureReg,\var{sm}) =      & \; (1,\pureReg,\var{sm}) \\
+  (0,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) =      & \; (0,\pureReg,\var{sm}) \\
 \end{align*}

### 2017-10-27 fully-abstract-comp/technical-report.tex
typos

@@
 \begin{align*}
-  (i,\normalReg,\var{sm}) \uplus (j,\normalReg,\var{sm}) =  & \; (\max(i,j),\normalReg,\var{sm}) \\
-  (1,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) =      & \; (1,\pureReg,\var{sm}) \\
-  (0,\pureReg,\var{sm}) \uplus (1,\pureReg,\var{sm}) =      & \; (1,\pureReg,\var{sm}) \\
-  (0,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) =      & \; (0,\pureReg,\var{sm}) \\
+  (i,\normalReg,\var{sm}) \oplus (j,\normalReg,\var{sm}) =  & \; (\max(i,j),\normalReg,\var{sm}) \\
+  (1,\pureReg,\var{sm}) \oplus (0,\pureReg,\var{sm}) =      & \; (1,\pureReg,\var{sm}) \\
+  (0,\pureReg,\var{sm}) \oplus (1,\pureReg,\var{sm}) =      & \; (1,\pureReg,\var{sm}) \\
+  (0,\pureReg,\var{sm}) \oplus (0,\pureReg,\var{sm}) =      & \; (0,\pureReg,\var{sm}) \\
 \end{align*}
 and for all other cases $\oplus$ is undefined.
-Specifically, $\uplus$ is not defined when both sides are an owned $\pureReg$-region.
+Specifically, $\oplus$ is not defined when both sides are an owned $\pureReg$-region.
 It is further not defined if the two sides do not agree on region type and state machine.

### 2017-10-30 fully-abstract-comp/technical-report.tex
moved ownership to pure regions only.

--- 2017-10-26 fully-abstract-comp/technical-report.tex
@@
 \newcommand{\Region}{\mathrm{Region}}
-\newcommand{\pureReg}{\mathrm{pure}}
-\newcommand{\normalReg}{\mathrm{normal}}
+\newcommand{\pure}{\mathrm{pure}}
+\newcommand{\pureo}{\mathrm{pure\_owned}}
+\newcommand{\normReg}{\mathrm{normal}}
 \newcommand{\State}{\mathrm{State}}
@@
 \[
-  \World = \RegionName \parfun 2 \times \Region
-\]
-where $\RegionName = \nats$
+  \World = \RegionName \parfun \Region
+\]
+where $\RegionName = \nats$.
 \[
@@
   \begin{array}{l}
-    \{\pureReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
-    \{\normalReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\pureo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\normReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
   \end{array} \right.
 \]
+where $\pure$ and $\pureo$ are regions governing segments of memory addressed by linear capabilities.
+$\pureo$ signifies that this region is addressable.
+$\pure$ signifies that the region is not owned and can thus not be addressed.
+Finally, $\normReg$ signifies that the region is only addressed by non-linear capabilities.
+Notice that no region allows for both linear and non-linear capabilities to address it.
+\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitely into the worlds (but maybe we can handle it in the memory satisfaction).}
 \lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
+
+Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\pureo$-regions.
 \[
@@
 \begin{align*}
-  (i,\normalReg,\var{sm}) \uplus (j,\normalReg,\var{sm}) = & (\max(i,j),\normalReg,\var{sm}) \\
-  (1,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) = & (1,\pureReg,\var{sm}) \\
-  (0,\pureReg,\var{sm}) \uplus (1,\pureReg,\var{sm}) = & (1,\pureReg,\var{sm}) \\
-  (0,\pureReg,\var{sm}) \uplus (0,\pureReg,\var{sm}) = & (0,\pureReg,\var{sm}) \\
+  (\normReg,\var{sm}) \oplus (\normReg,\var{sm}) =  & \; (\normReg,\var{sm}) \\
+  (\pureo,\var{sm}) \oplus (\pure,\var{sm}) = & \; (\pure,\var{sm}) \oplus (\pureo,\var{sm})\\
+                                           =  & \; (\pureo,\var{sm}) \\
+  (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
 \end{align*}
 and for all other cases $\oplus$ is undefined.
-Specifically, $\uplus$ is not defined when both sides are an owned $\pureReg$-region.
+Specifically, $\oplus$ is not defined when both sides are a $\pureo$-region.
 It is further not defined if the two sides do not agree on region type and state machine.
+Here $\var{sm}$ is the state machine of the region.
 

### 2017-10-30 fully-abstract-comp/technical-report.tex
small change in lr r
added a memory satisfaction definition.

@@
 \newcommand{\dom}{\mathrm{dom}}
+\newcommand{\powerset}[1]{\mathcal{P}(#1)}
 
@@
 \newcommand{\future}{\sqsupseteq}
-\newcommand{\privft}{\future^{\mathrm{priv}}}
-\newcommand{\pubft}{\future^{\mathrm{pub}}}
+\newcommand{\pub}{\mathrm{pub}}
+\newcommand{\privft}{\future^{\priv}}
+\newcommand{\pubft}{\future^{\pub}}
 
@@
 \lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
+  \[
+   \Rels = \{\phi_\pub \times \phi \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
+  \]
 
@@
 Here $\var{sm}$ is the state machine of the region.
+
+\subsection{Future world}
+
+\subsection{Memory satisfaction}
+\[
+  \memSat{\ms_S,\ms_\stk,\stk}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      \exists R_\ms : \dom(W) \fun \MemSeg^2 \ldotp \\
+      \quad \ms_T = \biguplus_{r \in \dom(W)} \pi_2(R_\ms(r)) \wedge \\
+      \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
+      \quad \exists r_0 , \dots, r_m, r_{m+1} \ldotp \pi_1(R_\ms(r_0,\dots,r_m)) = \ms_{\stk,0},\dots , \ms_{\stk,m} \wedge \\
+      \qquad \pi_1(R_\ms(r_{m+1})) = \ms_\stk \wedge \\
+      \qquad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
+      \qquad \exists R_w : \dom(W) \fun \World \ldotp\\
+      \quad \qquad W = \biguplus_{r \in \dom(W)} R_W(r) \wedge \\
+      \quad \qquad \forall r \in \dom(W) , n' < n \ldotp \\
+      \qquad \qquad \npair[n']{R(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_W (r))) 
+    \end{array}
+  \right.
+\]
+In the definition of memory satisfaction, $R_\ms$ acts as a partitioning in the two related memories.
+As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language.
+We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints impossed by the world are satisfied.
 
@@
       \quad S \text{ is a separation of $W$} \tand\\
-      \quad \forall r \in \dom(\reg_T) \setminus \{\pcreg, \rretd,\rretc \}\ldotp\\
+      \quad \forall r \in \RegName \setminus \{\pcreg, \rretd,\rretc \}\ldotp\\
       \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))

### 2017-11-01 fully-abstract-comp/technical-report.tex
typos and uplus to oplus for worlds
most of the permission based conditions to be used in value relation

@@
 \newcommand{\npair}[2][n]{\left(#1,#2\right)}
+
+\newcommand{\nsubeq}[1][n]{\overset{#1}{\subseteq}}
+\newcommand{\nsupeq}[1][n]{\overset{#1}{\supseteq}}
 
@@
 
-\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
-\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
+%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
+%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
+%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
+%\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
+%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
+\newcommand{\addressable}[1]{\plainfun{addressable}{#1}}
+\newcommand{\callCond}[1]{\plainfun{callCondition}{#1}}
+\newcommand{\decInstr}[1]{\plainfun{decodeInstruction}{#1}}
 \newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
 \newcommand{\encInstr}[1]{\plainfun{encodeInstruction}{#1}}
-\newcommand{\decInstr}[1]{\plainfun{decodeInstruction}{#1}}
+\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
+\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
+\newcommand{\exec}[1]{\plainfun{executable}{#1}}
+\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
+\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
+\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
+\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
+\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
+\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
+\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
+\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
+\newcommand{\range}[1]{\plainfun{range}{#1}}
+\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
+\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
+\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
+\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
 \newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
-%\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
-\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
-\newcommand{\exec}[1]{\plainfun{executable}{#1}}
-%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
-\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
+\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
 \newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
-%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
-%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
-\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
-\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
-%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
-\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
-\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
-\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
-\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
-\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
-\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
-\newcommand{\callCond}[1]{\plainfun{callCondition}{#1}}
-
-\newcommand{\range}[1]{\plainfun{range}{#1}}
 
@@
 \subsection{Worlds}
+\lau{30-10-2017: TODO: solve recursiveness in domain equation in the following. }
 \[
@@
 \[
-  W_1 \uplus W_1 = W
+  W_1 \oplus W_1 = W
   \begin{array}[t]{l}
@@
 \[
-  \memSat{\ms_S,\ms_\stk,\stk}{W} \text{ iff } 
+  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
   \left\{
@@
       \qquad \exists R_w : \dom(W) \fun \World \ldotp\\
-      \quad \qquad W = \biguplus_{r \in \dom(W)} R_W(r) \wedge \\
+      \quad \qquad W = \biguplus_{r \in \dom(W)} R_w(r) \wedge \\
       \quad \qquad \forall r \in \dom(W) , n' < n \ldotp \\
-      \qquad \qquad \npair[n']{R(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_W (r))) 
+      \qquad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
     \end{array}
@@
 We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints impossed by the world are satisfied.
+\lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
 
@@
 ,T},k_{c,T},k_{d,T}})\endarray\right)}
-      \in \lro(W\uplus W_\lrk \uplus W_\lrr \uplus W_\lrm)
+      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
     \end{array}
@@
                                           (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\dots,\dots})\endarray\right)}
-      \in \lro(W' \uplus W_\lrr \uplus W_\lrm)
+      \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
     \end{array}
@@
     \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W_\lrv) \ldotp\\
-      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \uplus W_\lrv)
+      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
     \end{array}
@@
     \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W_\lrv) \ldotp\\
-      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \uplus W_\lrv)
+      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
     \end{array}
@@
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \middle| \right)}
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \middle|
     \right\}
@@
 \]
+
+\subsection{Permission based conditions}
+\[
+  \addressable{\lin,W} =
+  \begin{cases}
+    \{ r \mid W(r) = (\normal,\_) \} & \text{if $\lin = \normal$} \\
+    \{ r \mid W(r) = (\pureo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
+  \end{cases}
+\]
+
+\[
+  \readCond{\lin,W} = \left\{ \npair{\baddr,\eaddr}) \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\lin, W} \ldotp \\
+      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
+      \qquad W(r) \nsubeq \iota_{[\baddr',\eaddr']}
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
+
+\lau{01-11-2017: TODO Write down address-stratified condition.}
+
+\[
+  \writeCond{\lin,W} = \left\{ \npair{\baddr,\eaddr}) \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\lin, W} \ldotp \\
+      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
+      \qquad W(r) \nsupeq[n-1] \iota_{[\baddr',\eaddr']} \wedge \\
+      \qquad W(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
+
+\[
+  \execCond{\lin,W} = \left\{ \npair{S_\perm,\baddr,\eaddr} \middle|
+    \begin{array}{l}
+      \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
+      \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{(c_{d,S},c_{d,T})} \in \lrv(W'') \wedge \\
+      \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lrp(W' \oplus W'') \Rightarrow \\
+      \quad \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lre(W' \oplus W'')
+    \end{array}
+    \right\}
+\]
+
+\subsection{Standard regions}
+\label{sec:standard-regions}
+Standard region:
+\[
+  \iota_A \defeq (*,=,=,H_A)
+\]
+where $H_A$ is defined as follows:
+\[
+  H_A \; s \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
+    \begin{array}{l}
+      \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
+      \forall \aaddr \in A \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(\xi(\hat{W}))
+    \end{array}
+  \right\}
+\]
+\lau{01-11-2017: TODO look into what kind of world separation should happen here.}
 

### 2017-11-03 fully-abstract-comp/technical-report.tex
typos and minor omissions.

@@
 \begin{align*}
-  \tsem{\tsplit{r_1}{r_2}{r_3}{r_4}} = &
+  \sem{\tsplit{r_1}{r_2}{r_3}{r_4}} = &
                                \begin{cases}
@@
 I think we talked about this, and it is something that should be handled like linking (a fresh seal is available in the beginning of a programs code along with a capability for a linking table)).
-    % restrict the stack capability to the unused part (and clear this part).
+    % restrict the stack capability to the unused part (and clear this part (is this really necessary?)).
     % make return capability and ceal it:
@@
     % restore code:
+    %   check the stack end address.
     %   move the stack pointer to the proper register
@@
 \lau{30-10-2017: TODO: solve recursiveness in domain equation in the following. }
+\lau{02-11-2017: We may need revoked regions again to handle the stack.}
 \[
@@
 \[
-  \iota_A \defeq (*,=,=,H_A)
-\]
+  \iota_A \defeq (\normal,*,=,=,H_A)
+\]
+\lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
 where $H_A$ is defined as follows:

### 2017-11-03 fully-abstract-comp/technical-report.tex
made it more clear what can be both registers and numbers in the op. sem.
fleshed out some details in continuation relation

@@
   & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
-      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
+      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
 \end{array}
@@
 \begin{align*}
-  \sem{\tjnz{r_1}{r_2}}(\Phi) = &       
+  \sem{\tjnz{r}{\rn}}(\Phi) = &       
                              \begin{cases}
@@
                                \begin{array}[t]{rl}
-                                 \Phi&\updReg{r_1}{w}\\
-                                     &\updReg{\pcreg}{\Phi(r_1)}
-                               \end{array}
-                                               & w = \linCons{\Phi(r_1)} \tand \nonZero{\Phi(r_2)}\\
+                                 \Phi&\updReg{r}{w}\\
+                                     &\updReg{\pcreg}{\Phi(r)}
+                               \end{array}
+                                               & w = \linCons{\Phi(r)} \tand \nonZero{\Phi(\rn)}\\
                                \updPcAddr{\Phi} & \totherwise
@@
 \begin{align*}
-  \sem{\tmove{r_1}{r_2}}(\Phi) = & 
-                              \begin{cases}
-                                \updPcAddr{\Phi\updReg{r_1}{r_2}} & r_2 \in \ints \\
-                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
-                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
-                              \end{cases}
-\end{align*}
-Notice that in the case where we are moving a linear capability and $r_1 = r_2$ the order of the updates matter.
+  \sem{\tmove{r}{\rn}}(\Phi) = & 
+                              \begin{cases}
+                                \updPcAddr{\Phi\updReg{r}{\rn}} & \rn \in \ints \\
+                                \updPcAddr{\Phi\updReg{\rn}{w}\updReg{r}{\Phi(\rn)}} & w = \linCons{\Phi(\rn)}
+                              \end{cases}
+\end{align*}
+Notice that in the case where we are moving a linear capability and $r = \rn$ the order of the updates matter.
 
@@
 \begin{align*}
-  \sem{\tcca{r_1}{r_2}}(\Phi) = & 
-                                  \begin{cases}
-                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
-                                    \arraycolsep=0pt
-                                    \begin{array}[t]{l}
-                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+  \sem{\tcca{r}{\rn}}(\Phi) = & 
+                                  \begin{cases}
+                                    \updPcAddr{\Phi\updReg{r}{c}} &  
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
                                       \quad\text{ and in either case $n \in \ints$ and} \\
-                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                      \Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                       c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
@@
                                     \\
-                                    \updPcAddr{\Phi\updReg{r_1}{c}} &  
-                                    \arraycolsep=0pt
-                                    \begin{array}[t]{l}
-                                      \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                    \updPcAddr{\Phi\updReg{r}{c}} &  
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
                                       \quad\text{ and in either case $n \in \ints$ and} \\
-                                      \Phi(r_1) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
+                                      \Phi(r) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
                                       c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
@@
                                     \\
-                                    \src{\updPcAddr{\Phi\updReg{r_1}{c}}} &  
+                                    \src{\updPcAddr{\Phi\updReg{r}{c}}} &  
                                     \sourcecolor\left.
@@
                                     \begin{array}[t]{l}
-                                      \text{either $n = r_2$ or $n = \Phi(r_2)$}\\
+                                      \text{either $n = \rn$ or $n = \Phi(\rn)$}\\
                                       \quad\text{ and in either case $n \in \ints$ and} \\
-                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                      \Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                       c = \stkptr{\perm,\baddr,\eaddr,\aaddr + n}
@@
 \begin{align*}
-  \sem{\trestrict{r_1}{r_2}{r_3}} = &
+  \sem{\trestrict{r_1}{r_2}{\rn}} = &
                                       \begin{cases}
@@
                                           \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                          \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
                                           \quad\text{and in either case $n \in \ints$ and} \\
@@
                                           \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
+                                          \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
                                           \quad\text{and in either case $n \in \ints$ and} \\
@@
 \begin{align*}
-  \sem{\tplus{r_0}{r_1}{r_2}}(\Phi) = &
+  \sem{\tplus{r_0}{\rn_1}{\rn_2}}(\Phi) = &
                                                   \begin{cases}
@@
                                                       \text{if for $i \in \{1,2\}$}\\
-                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
+                                                      \quad\text{$n_i = \rn_i$ or $n_i = \Phi(\rn_i)$}\\
                                                       \quad\text{and in either case $n_i \in \ints$}\\
@@
 \begin{align*}
-  \sem{\tminus{r_0}{r_1}{r_2}}(\Phi) = &
+  \sem{\tminus{r_0}{\rn_1}{\rn_2}}(\Phi) = &
                                                   \begin{cases}
@@
                                          \Phi(r_1) = \retptrc(\sigma,\_) \tand \\
-                                         \Phi(r_2) = \retptrd(\sigma,\_) \tand \\
+                                         \Phi(r_2) = \retptrd(\sigma,(\aaddr_\stk,\_)) \tand \\
                                          \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
                                          \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
-                                         \stkf = (\opc,\ms_{\stk,\priv},\aaddr_\stk) \tand \\
+                                         \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
                                          \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
@@
 \begin{align*}
-  \sem{\tsplit{r_1}{r_2}{r_3}{r_4}} = &
+  \sem{\tsplit{r_1}{r_2}{r_3}{\rn_4}} = &
                                \begin{cases}
@@
                                    \Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                   \Phi(r_4) = n \tor r_4 = n \\
+                                   \Phi(\rn_4) = n \tor \rn_4 = n \\
                                    \quad\text{ and in either case } n \in \nats \\
@@
                                    \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
-                                   \Phi(r_4) = n \tor r_4 = n \\
+                                   \Phi(\rn_4) = n \tor \rn_4 = n \\
                                    \quad\text{ and in either case } n \in \nats \\
@@
                                    \Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                   \Phi(r_4) = n \tor r_4 = n \\
+                                   \Phi(\rn_4) = n \tor \rn_4 = n \\
                                    \quad\text{ and in either case } n \in \nats \\
@@
 
-\lau{19-10-2017: $\lrk$ may indicate that $\lre$ should really be a relation between a \src{source pc} and a pair of a target pc and stack pointer?
-Otherwise, what happens with the data part of the return pointer?}
-\[
-  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,\_)},\\\src{\retptrc(\sigma,\_)},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
+\lau{03-11-2017: TODO take a look at linearity of this definition - in paticular the recovered pc and stack pointer.}
+\[
+  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
     \begin{array}{l}
-      k_{d,T} \text{ is a stack pointer with correct end and such with everything related to the source stack} \\
-      k_{c,T} \text{ relates to the old pc on the stack}\\
+      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
+      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
+      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
+      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
       \forall n' \leq n, W' \pubft W \ldotp \\
       \quad \forall, W_\lrr, W_\lrm \ldotp  \\
+      \qquad\opc = c_c \wedge\\
+      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
       \qquad\npair[n']{\src{\retptrd(\sigma,\_)},\src{\retptrc(\sigma,\_)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \qquad\dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk] \wedge \\
+      \qquad\eaddr_\stk = \aaddr_d-1\\
       \qquad\Rightarrow\\
-      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\dots}},\stk, \ms_\stk)},\\
-                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\dots,\dots})\endarray\right)}
+      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\_}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
+                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\_)})\endarray\right)}
       \in \lro(W' \oplus W_\lrr \oplus W_\lrm)

### 2017-11-03 fully-abstract-comp/technical-report.tex
added a partial implementation for call. Need to decide what we do about seals.

@@
 \lau{16-10-2017: Implementation of $\scall{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
+\[
+  \begin{array}{l}
+    \tgete{r_{1t}}{\rstk}\\
+    \tstore{\rstk}{r_{t1}}\\
+    \tcca{\rstk}{-1}\\
+    \tgeta{r_{t1}}{\rstk}\\
+    \tsplit{\rstk}{\rretd}{\rstk}{r_{t1}}\\
+    \tmove{r_{t1}}{\pcreg}\\
+    \tsetatob{r_{t1}}{r_{t1}}\\
+    \tload{r_{t1}}{r_{t1}}\\
+    \tcca{r_{t1}}{{\color{red} \dots}}\\
+    \tcseal{\rretd}{\rretd}{r_{t1}}\\
+    \tmove{\rretc}{\pcreg}\\
+    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
+    \tcseal{\rretc}{\rretc}{r_{t1}}\\
+    \tmove{r_{t1}}{0}\\
+    \txjmp{r_1}{r_2}\\
+    \text{// the following is the return code}\\
+    \tgetb{r_{t1}}{\rstk}\\
+    \tminus{r_{t1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
+    \tmove{r_{t2}}{\pcreg}\\
+    \tcca{r_{t2}}{5} \text{ //magic number is the offset to fail} \\
+    \tjnz{r_{t2}}{r_{t1}} \\
+    \tcca{r_{t2}}{1} \text{ //magic number is the offset to after fail} \\
+    \tjmp{r_{t2}} \\
+    \tfail \\
+    \tsplice{\rstk}{\rstk}{\rdata} \\
+    \tcca{\rstk}{1}\\
+    \tmove{r_{t1}}{0}\\
+    \tmove{r_{t2}}{0}
+  \end{array}
+\]
+{\color{red} \dots} needs to be the offset to the seal we want to use.
 

### 2017-11-06 fully-abstract-comp/technical-report.tex
added description of call implementation and a comment about future worlds and revoked regions (should we need them).

@@
 \usepackage{stmaryrd}
+\usepackage{mathpartir}
 
@@
 {\color{red} \dots} needs to be the offset to the seal we want to use.
+The call code does the following:
+\begin{itemize}
+\item Store the end address of the stack to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
+\item Get the current address of the stack pointer and split according to it.
+\item Retrieve the seal of the program.
+\item cca the seal, so the seal to be used is active 
+\item seal our private part of the stack capability.
+\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
+\item clear the temporaryu register.
+\item cross jump to the two specified registers.
+\item Upon return:
+  \begin{itemize}
+  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
+  \item splice the returned stack pointer and the stack pointer for our private stack.
+  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
+  \item clear the temporary registers.
+  \end{itemize}
+\end{itemize}
 
@@
 \lau{02-11-2017: We may need revoked regions again to handle the stack.}
+\lau{06-11-2017: If we need to introduce revoked regions like we hade for the local capabilities, we may consider the following: For the future world relation require the existance of a bijection from region names to region names.
+For all regions in the present world the future region is given by the bijection.
+This would mean that reinstantiation would happen simply by giving a bijection that maps to the new region.
+In other words, the weird shadowing would not be necessary.
+  \begin{mathpar}
+    \inferrule{ \dom(W') \supseteq \dom(W) \\ \exists m : \RegionName \fun \RegionName, \text{ $m$ injective} \ldotp \forall r \in \dom(W) \ldotp W'(m(r)) \pubft W(r) }
+    { W' \pubft W }
+  \end{mathpar}
+}
 \[

### 2017-11-07 fully-abstract-comp/technical-report.tex
added address stratified definition
added the 'normal' capabilities to the value relation
Typo.
typo.
renamed regions (pure to spatial, normal to pure)

@@
 \newcommand{\Region}{\mathrm{Region}}
+\newcommand{\spatial}{\mathrm{spatial}}
+\newcommand{\spatialo}{\mathrm{spatial\_owned}}
 \newcommand{\pure}{\mathrm{pure}}
-\newcommand{\pureo}{\mathrm{pure\_owned}}
-\newcommand{\normReg}{\mathrm{normal}}
 \newcommand{\State}{\mathrm{State}}
@@
   \begin{array}{l}
-    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
-    \{\pureo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
-    \{\normReg \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
+    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
+    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
   \end{array} \right.
 \]
-where $\pure$ and $\pureo$ are regions governing segments of memory addressed by linear capabilities.
-$\pureo$ signifies that this region is addressable.
-$\pure$ signifies that the region is not owned and can thus not be addressed.
-Finally, $\normReg$ signifies that the region is only addressed by non-linear capabilities.
+where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities.
+$\spatialo$ signifies that this region is addressable.
+$\spatial$ signifies that the region is not owned and can thus not be addressed.
+Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities.
 Notice that no region allows for both linear and non-linear capabilities to address it.
@@
 
-Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\pureo$-regions.
+Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions.
 \[
@@
 \begin{align*}
-  (\normReg,\var{sm}) \oplus (\normReg,\var{sm}) =  & \; (\normReg,\var{sm}) \\
-  (\pureo,\var{sm}) \oplus (\pure,\var{sm}) = & \; (\pure,\var{sm}) \oplus (\pureo,\var{sm})\\
-                                           =  & \; (\pureo,\var{sm}) \\
   (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
+  (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
+                                           =  & \; (\spatialo,\var{sm}) \\
+  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
 \end{align*}
 and for all other cases $\oplus$ is undefined.
-Specifically, $\oplus$ is not defined when both sides are a $\pureo$-region.
+Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region.
 It is further not defined if the two sides do not agree on region type and state machine.
@@
       \qquad \exists R_w : \dom(W) \fun \World \ldotp\\
-      \quad \qquad W = \biguplus_{r \in \dom(W)} R_w(r) \wedge \\
+      \quad \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
       \quad \qquad \forall r \in \dom(W) , n' < n \ldotp \\
@@
 \lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out.
-I can see how it will enforce linearity, but is you give a "pure" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part.
+I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part.
 In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file.
@@
 \]
-Where $S\text{ is a separation of $W$}$ means that no two partion share the same ``pure'' region (whatever that is).
+Where $S\text{ is a separation of $W$}$ means that no two partion share the same ``spatial'' region (whatever that is).
 
@@
   \begin{array}[t]{l}
-    \left\{ \npair{\stpair[.]{i}{i}} \middle| i \in \ints \right\}\cup \\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\_)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \middle| 
+    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\_)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
     \begin{array}{l}
@@
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\_)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \middle| 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\_)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \;\middle| \;
     \begin{array}{l}
@@
     \right\}\cup \\
-    \left\{ \npair{\left(\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \middle|
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
     \begin{array}{l}
@@
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \middle|
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} \;\middle|\;
+    \right\} \cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \right\} \cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \begin{array}{r l l }
+      \perm \in \readAllowed{} &\Rightarrow& \npair{\baddr,\eaddr} \in \readCond{\lin,W}\\
+      \perm \in \writeAllowed{} &\Rightarrow& \npair{\baddr,\eaddr} \in \writeCond{\lin,W}\\
+      \perm = \rwx &\Rightarrow& \npair{\{\rwx,\rx\},\baddr,\eaddr} \in \execCond{\lin,W}\\
+      \perm = \rx &\Rightarrow& \npair{\{\rx\},\baddr,\eaddr} \in \execCond{\lin,W}
+    \end{array}
+
     \right\}
@@
     \{ r \mid W(r) = (\normal,\_) \} & \text{if $\lin = \normal$} \\
-    \{ r \mid W(r) = (\pureo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
+    \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
   \end{cases}
@@
 
-\lau{01-11-2017: TODO Write down address-stratified condition.}
+\begin{definition}
+  \label{def:address-stratified}
+  We say that a region $\iota = (\_,\_,\_,\_,H)$ is address stratified iff
+  \[
+    \begin{array}{l}
+      \forall n, \src{\ms_S},\ms_T,\src{\ms_S'},\ms_T',s,\hat{W}\ldotp \\
+      \quad \npair{\stpair{\ms}{\ms}}, \npair{\stpair[.]{\ms_S'}{\ms_T'}} \in H \; s \; \hat{W} \wedge \\
+      \quad \dom(\src{\ms_S}) = \dom(\ms_T) = \dom(\src{\ms_S'}) = \dom(\ms_T') \\
+      \quad \Rightarrow \\
+      \qquad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\src{\ms_S}\update{\aaddr}{\src{\ms_S'}(\aaddr)},\ms_T\update{\aaddr}{\ms_T'(\aaddr)})} \in H \; s \; \hat{W}
+    \end{array}
+  \]
+\end{definition}
 

### 2017-11-07 fully-abstract-comp/technical-report.tex
corrected where the values from the stack pointer comes from.
Typos

@@
 \dominique{27-09-2017: I think we need to do something about tail calls in the source language?}
-\lau{27-09-2017: In what sense? Do you want to allow them (I do not thinkn they are allowed at the moment?)}
+\lau{27-09-2017: In what sense? Do you want to allow them (I do not think they are allowed at the moment?)}
 
@@
 The $\encInstr{}$ function must be injective.
-Futher $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
+Further $\decInstr{}$ must be the left inverse of $\encInstr{}$ that is for all $i \in \Instr$
 \[
@@
 \item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
-\item clear the temporaryu register.
+\item clear the temporary register.
 \item cross jump to the two specified registers.
@@
   A register file $\reg$ is said to respect write-xor-execute with slack $[\bar{r}]$ iff \\
-  For all $r \in RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
+  For all $r \in \RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
 ~\\
@@
 \lau{02-11-2017: We may need revoked regions again to handle the stack.}
-\lau{06-11-2017: If we need to introduce revoked regions like we hade for the local capabilities, we may consider the following: For the future world relation require the existance of a bijection from region names to region names.
+\lau{06-11-2017: If we need to introduce revoked regions like we had for the local capabilities, we may consider the following: For the future world relation require the existence of a bijection from region names to region names.
 For all regions in the present world the future region is given by the bijection.
@@
 Notice that no region allows for both linear and non-linear capabilities to address it.
-\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitely into the worlds (but maybe we can handle it in the memory satisfaction).}
+\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
 \lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
@@
 \[
-  W_1 \oplus W_1 = W
+  W_1 \oplus W_2 = W
   \begin{array}[t]{l}
@@
 As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language.
-We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints impossed by the world are satisfied.
+We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints imposed by the world are satisfied.
 \lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
@@
 \lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
-\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem staisfaction).
+\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem satisfaction).
 That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
@@
 \lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out.
-I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfation does not have it, and it cannot find a memory segment for this part.
+I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part.
 In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file.
@@
 
-\lau{03-11-2017: TODO take a look at linearity of this definition - in paticular the recovered pc and stack pointer.}
+\lau{03-11-2017: TODO take a look at linearity of this definition - in particular the recovered pc and stack pointer.}
 \[
@@
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
-      \qquad\dom(\ms_\stk) = [\baddr_\stk,\eaddr_\stk] \wedge \\
+      \qquad\src{reg_S}(\rstk) = \src{\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_}} \wedge \\
       \qquad\eaddr_\stk = \aaddr_d-1\\
       \qquad\Rightarrow\\
-      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\_}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
-                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\_)})\endarray\right)}
+      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\aaddr_d}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
+                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\aaddr_d)})\endarray\right)}
       \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
@@
       \quad S \text{ is a separation of $W$} \tand\\
-      \quad \forall r \in \RegName \setminus \{\pcreg, \rretd,\rretc \}\ldotp\\
+      \quad \forall r \in \RegName \setminus \{\pcreg, \rretd,\rretc,\rdata \}\ldotp\\
       \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
@@
 \]
-Where $S\text{ is a separation of $W$}$ means that no two partion share the same ``spatial'' region (whatever that is).
+Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
 
@@
 We basically wanted a formalism in which we could design one region that would govern the stack.
-I do not remeber why it was made possible with PORs.}
+I do not remember why it was made possible with PORs.}
   \begin{itemize}
@@
 
-One way to solve this problem is to make the "top addess" of the stack known. There are many ways to do this, but we have chosen to do the following:
+One way to solve this problem is to make the "top address" of the stack known. There are many ways to do this, but we have chosen to do the following:
 \begin{itemize}

### 2017-11-08 fully-abstract-comp/technical-report.tex
comment
Remove some obsolete notes, add a new comment.

@@
 
-\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine.
-I could not recall why we had this, so I removed it.
-(I think it may have been from when we had call and return instructions?)}
-\dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
-\lau{28-09-2017: I can see from the commit history that this was something when we had push/pop instructions as well as a return instruction. I have removed the restriction on $\rstk$.}
-
 \subsection{Permissions}
@@
 \lau{27-09-2017: In what sense? Do you want to allow them (I do not think they are allowed at the moment?)}
+\dominique{8-11-2017: Yes, I would like to support them.
+I think it would be a strength of our calling convention that we can support them (as I expect we can) and it would be good to demonstrate this.}
 

### 2017-11-08 fully-abstract-comp/technical-report.tex
Add comments to implementation of call and add a note.

@@
   \begin{array}{l}
-    \tgete{r_{1t}}{\rstk}\\
+    \text{// push current end address of stack pointer on the stack.}\\
+    \tgete{r_{t1}}{\rstk}\\
     \tstore{\rstk}{r_{t1}}\\
     \tcca{\rstk}{-1}\\
+    \text{// split the stack at its current address.}\\
     \tgeta{r_{t1}}{\rstk}\\
     \tsplit{\rstk}{\rretd}{\rstk}{r_{t1}}\\
+    \text{// load the seal for the return pointer through the pc capability.}\\
     \tmove{r_{t1}}{\pcreg}\\
@@
     \tcca{r_{t1}}{{\color{red} \dots}}\\
+    // \text{seal the used stack frame as the data part of the return pointer pair.}\\
     \tcseal{\rretd}{\rretd}{r_{t1}}\\
+    // \text{obtain the code part of the return pointer pair and seal it too.}\\
     \tmove{\rretc}{\pcreg}\\
@@
     \tcseal{\rretc}{\rretc}{r_{t1}}\\
+    // \text{now clear temporary register and jump to the adversary.}\\
     \tmove{r_{t1}}{0}\\
@@
     \text{// the following is the return code}\\
+    \text{// check that the stack pointer is the same we handed out.}\\
     \tgetb{r_{t1}}{\rstk}\\
@@
     \tfail \\
+    \text{// join our stored private stack frame with the rest of the stack}\\
     \tsplice{\rstk}{\rstk}{\rdata} \\
+    \text{// pop the stored stack end address}\\
     \tcca{\rstk}{1}\\
+    \text{// clear temporary registers used}\\
     \tmove{r_{t1}}{0}\\
-    \tmove{r_{t2}}{0}
-  \end{array}
-\]
+    \tmove{r_{t2}}{0}\\
+    \text{// continue program after invocation.}
+  \end{array}
+\]
+\dominique{Return code could be simplified if we had jz in addition to jnz.}
+\dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
+  I understand this is for ensuring that our private stack frame is not empty, which is important.
+  But why don't we just push a constant then?
+  Also: perhaps we could only do it if the current private stack frame is empty to begin with?
+}
 {\color{red} \dots} needs to be the offset to the seal we want to use.

### 2017-11-08 fully-abstract-comp/technical-report.tex
merge

@@
   \begin{array}{l}
-    \text{// push current end address of stack pointer on the stack.}\\
-    \tgete{r_{t1}}{\rstk}\\
-    \tstore{\rstk}{r_{t1}}\\
+    \text{// push 42 on the stack (so it is non-empty).}\\
+    \tstore{\rstk}{42}\\
     \tcca{\rstk}{-1}\\
@@
       \forall n' \leq n, W' \pubft W \ldotp \\
-      \quad \forall, W_\lrr, W_\lrm \ldotp  \\
+      \quad \forall W_\lrr, W_\lrm \ldotp  \\
       \qquad\opc = c_c \wedge\\
@@
     \begin{array}{r l l }
-      \perm \in \readAllowed{} &\Rightarrow& \npair{\baddr,\eaddr} \in \readCond{\lin,W}\\
-      \perm \in \writeAllowed{} &\Rightarrow& \npair{\baddr,\eaddr} \in \writeCond{\lin,W}\\
-      \perm = \rwx &\Rightarrow& \npair{\{\rwx,\rx\},\baddr,\eaddr} \in \execCond{\lin,W}\\
-      \perm = \rx &\Rightarrow& \npair{\{\rx\},\baddr,\eaddr} \in \execCond{\lin,W}
+      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
+      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
+      \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
+      \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
     \end{array}
@@
 \[
-  \readCond{\lin,W} = \left\{ \npair{\baddr,\eaddr}) \middle| 
+  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
     \begin{array}{l}
@@
 \[
-  \writeCond{\lin,W} = \left\{ \npair{\baddr,\eaddr}) \middle| 
+  \writeCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
     \begin{array}{l}
@@
 \[
-  \execCond{\lin,W} = \left\{ \npair{S_\perm,\baddr,\eaddr} \middle|
+  \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
     \begin{array}{l}
@@
 
+Eventhough we call them linear, they are more like \emph{affine} capabilities because our current operational semantics allows for them to be overwritten.
+We choose to stick with the name as the term linear is also being misused by others for instance in Rust.
+
 \subsubsection{Write-linear capabilities}

### 2017-11-08 fully-abstract-comp/technical-report.tex
changed call impl to use magic number. Updated comments.

@@
     \tcca{\rstk}{-1}\\
-    \text{// split the stack at its current address.}\\
+    \text{// split the stack at its current address - rstk done.}\\
     \tgeta{r_{t1}}{\rstk}\\
@@
     \tcca{r_{t1}}{{\color{red} \dots}}\\
-    // \text{seal the used stack frame as the data part of the return pointer pair.}\\
+    \text{// seal the used stack frame as the data part of the return pointer pair.}\\
     \tcseal{\rretd}{\rretd}{r_{t1}}\\
-    // \text{obtain the code part of the return pointer pair and seal it too.}\\
+    \text{// obtain the code part of the return pointer pair and seal it too.}\\
     \tmove{\rretc}{\pcreg}\\
@@
     \tcseal{\rretc}{\rretc}{r_{t1}}\\
-    // \text{now clear temporary register and jump to the adversary.}\\
+    \text{// now clear temporary register and jump to the adversary.}\\
     \tmove{r_{t1}}{0}\\
@@
     \tfail \\
-    \text{// join our stored private stack frame with the rest of the stack}\\
+    \text{// join our stored private stack frame with the rest of the stack (this also finishes the stack pointer check).}\\
     \tsplice{\rstk}{\rstk}{\rdata} \\
-    \text{// pop the stored stack end address}\\
+    \text{// pop the magic number 42}\\
     \tcca{\rstk}{1}\\

### 2017-11-08 fully-abstract-comp/technical-report.tex
Added a new argument to call, namely the offset in the seal that it uses.

@@
 % scall - special two instruction
-\newcommand{\scall}[3][]{  
-\ifthenelse{\equal{#2#3}{}}
-  {\ensuremath{\zinstr{\src{call}}_{#1}}}
-  {\ensuremath{\zinstr{\src{call}}_{#1} \; #2 \; #3}}
+\newcommand{\scall}[4][]{  
+\ifthenelse{\equal{#3#4}{}}
+  {\ensuremath{\zinstr{\src{call}}_{#1}^{#2}}}
+  {\ensuremath{\zinstr{\src{call}}_{#1}^{#2} \; #3 \; #4}}
 }
@@
     n & \in & \nats \\
+    o & \in & \nats \\
     \src{r} &\in &  \RegName\\
     \src{\rn} &\defbnf & \src{r} \mid n \\
-    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
-  \end{array}
-\]
-There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.
-
-The source machine allows for variable length instructions which we utilize for $\scall{}{}$.
-In other words, when we $\scall{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction.
+    \src{i} & \defbnf &  \trg{i} \mid \scall{o}{\src{r}}{\src{r}}
+  \end{array}
+\]
+There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{o}{}{}$ instruction (the $o$ refers to the seal it uses, namely the offset in the seals made available by linking).
+
+The source machine allows for variable length instructions which we utilize for $\scall{o}{}{}$.
+In other words, when we $\scall{o}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction.
 This is described in more detail when we present the decoding function.
@@
 \paragraph{Step relation}
-The instruction $\scall{}{}$ has length $\calllen$ which means that it does not fit on one memory address.
-In fact, $\scall{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
-We therefore introduce $\scall[1]{r_1}{r_2},\dots,\scall[\calllen-1]{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
-We define the following condition that indicates that $\aaddr$ is the first address of a $\scall{r_1}{r_2}$ instruction in the configuration $\Phi$:
+The instruction $\scall{o}{}{}$ has length $\calllen$ which means that it does not fit on one memory address.
+In fact, $\scall{o}{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
+We therefore introduce $\scall[1]{o}{r_1}{r_2},\dots,\scall[\calllen-1]{o}{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{o}{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
+We define the following condition that indicates that $\aaddr$ is the first address of a $\scall{o}{r_1}{r_2}$ instruction in the configuration $\Phi$:
 \[
@@
       \begin{array}{l}
-        \Phi.\mem(\aaddr) = \scall[1]{r_1}{r_2} \tand\\
+        \Phi.\mem(\aaddr) = \scall[1]{o}{r_1}{r_2} \tand\\
         \vdots \\
-        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen]{r_1}{r_2}
+        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen]{o}{r_1}{r_2}
       \end{array}
@@
 \begin{align*}
-  \src{\Phi} & \; \src{\step \sem{\scall{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
+  \src{\Phi} & \; \src{\step \sem{\scall{o}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                   \begin{array}[t]{l}
@@
 \label{par:call-impl}
-\lau{16-10-2017: Fill this paragraph with $\scall{}{}$ implementation, i.e., the instructions $\scall[1]{r_1}{r_2} \dots \scall[\calllen]{r_1}{r_2}$ is equal to.
-Add convenience notation like $\update{\mem.a}{\scall{r_1}{r_2}} = \dots$}
-\lau{16-10-2017: Implementation of $\scall{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
+\lau{16-10-2017: Fill this paragraph with $\scall{o}{}{}$ implementation, i.e., the instructions $\scall[1]{o}{r_1}{r_2} \dots \scall[\calllen]{o}{r_1}{r_2}$ is equal to.
+Add convenience notation like $\update{\mem.a}{\scall{o}{r_1}{r_2}} = \dots$}
+\lau{16-10-2017: Implementation of $\scall{o}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
 \[
@@
     \tload{r_{t1}}{r_{t1}}\\
-    \tcca{r_{t1}}{{\color{red} \dots}}\\
+    \tcca{r_{t1}}{o}\text{ // $o$ is the offset this call uses.}\\
     \text{// seal the used stack frame as the data part of the return pointer pair.}\\
@@
 }
-{\color{red} \dots} needs to be the offset to the seal we want to use.
 The call code does the following:
@@
 \begin{align*}
-  \sem{\scall{\src{r_1}}{\src{r_2}}}(\Phi) = & 
+  \sem{\scall{o}{\src{r_1}}{\src{r_2}}}(\Phi) = & 
                                                \begin{cases}
@@
 \[
-  \comp{\scall{r_1}{r_2}} = 
+  \comp{\scall{o}{r_1}{r_2}} = 
   \begin{array}[t]{l}
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\scall{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2'}$ and there exists no $j<i$ s.t.
+  and $\src{\Phi'}$ points to $\src{\scall{o}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2'}$ and there exists no $j<i$ s.t.
   \[
@@
   \]
-  where $\src{\Phi''}$ points to $\src{\scall{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
+  where $\src{\Phi''}$ points to $\src{\scall{o}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
   then
@@
 That said, I would like to keep it around (maybe change it so it becomes an argument for variable length instructions).}
-This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{}{}$ and $\sreturn$ instructions.
+This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{o}{}{}$ and $\sreturn$ instructions.
 The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.
@@
 \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-$\scall{r5}{0}$
+$\scall{o}{r5}{0}$
 $\sjmp{r5}$
 \end{lstlisting}
-\footnote{At this time $\scall{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
+\footnote{At this time $\scall{o}{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
 We would not have enough instructions to set up a proper return pointer for the compiled return to use.
@@
 We do not offer primitive \sreturn{}, \spush{} and \spop{} instructions, even though there is nothing keeping us from doing that, either as an additional primitive instruction or as a macro on top of the primitive token values mentioned above.
-We do offer a primitive \scall{}{} instruction, and it offers the only way to create a new stack frame.
+We do offer a primitive \scall{o}{}{} instruction, and it offers the only way to create a new stack frame.
 It is the only instruction that does not have a target language counterpart.
-However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{}{} compiles to.
+However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{o}{}{} compiles to.
 We will discuss later that this means that if we back translate an adversary piece of code (in the full abstraction proof), then it will be back translated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
@@
 % Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
-% \lau{07-09-2017: One would use $\scall{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
+% \lau{07-09-2017: One would use $\scall{o}{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
 The return would happen just by using $\sxjmp{}{}$ with the $\retptrd$ and $\retptrc$.
@@
 In other words, we need something to recognise when a return happens.}
-% \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{}{}$ instruction.
+% \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{o}{}{}$ instruction.
 The real implementation of the $\retptrc$ will simply point to a piece of clean up code that splices together the old/private stack pointer with the one we got from the untrusted party.
-This may, however, give some issues when we want to reason about it because there will be a disconnect between where the clean up job is done in the operational semantics of the source language (namely as part of the $\sxjmp{}{}$), but in a translated program, the code that does the clean up origins with the $\scall{}{}$ instruction.
+This may, however, give some issues when we want to reason about it because there will be a disconnect between where the clean up job is done in the operational semantics of the source language (namely as part of the $\sxjmp{}{}$), but in a translated program, the code that does the clean up origins with the $\scall{o}{}{}$ instruction.
 }
@@
 The reason is that we only want to prove a result about programs that respect the calling convention, i.e.\ that do not do anything stupid.
-For example, in the old setup with global/local capabilities, when trusted code performs an \scall{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
+For example, in the old setup with global/local capabilities, when trusted code performs an \scall{o}{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
 Such sanity conditions are necessary (we cannot guarantee anything for code that does such stupid things) but we do not want to spend computation time on checking them.

### 2017-11-15 fully-abstract-comp/technical-report.tex
Refined worlds and priv. future world relation

@@
 \newcommand{\pure}{\mathrm{pure}}
+\newcommand{\revoked}{\mathrm{revoked}}
 \newcommand{\State}{\mathrm{State}}
@@
 }
+\lau{10-11-2017: The above may not work well here because we do not want a $\spatialo$ region to be taken over by a $\pure$ region (as this would allow for normal capabilities addressing an area of memory that was only addressable by linear capabilities before).}
 \[
@@
   \begin{array}{l}
-    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
-    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\\
-    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})
+    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\uplus \\
+    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2}) \uplus\\
+    \{\revoked\}
   \end{array} \right.
@@
 \subsection{Future world}
+
+For both $\privft$ and $\pubft$ we have
+\begin{mathpar}
+  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
+            { W' \future W }
+\end{mathpar}
+
+The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
+\begin{mathpar}
+  \inferrule{ }
+            { \revoked \privft (\spatial,\_,\_,\_,\_)}
+\and
+  \inferrule{ }
+            { \revoked \privft \revoked }
+\and
+  \inferrule{ (s,s') \in \phi }
+            { (v,s',\phi^\pub,\phi,H) \privft (v,s,\phi^\pub,\phi,H)  }
+
+\end{mathpar}
 

### 2017-11-15 local-capability-machine/local-capability-machine.tex
Added comment about macros.
added first draft for the days discussions
typo in lt impl
typo
updates to the local cap. tech report

@@
                  \jmp{\lv} \mid 
-                 \jnz{\lv}{\rv} \mid
+                 \jnz{\lv}{\lv} \mid
                  \move{\lv}{\rv} \mid 
@@
                  \getp{\lv}{\lv} \mid
+
                  \getl{\lv}{\lv} \mid 
@@
 \begin{align*}
-  \Phi & \rightarrow \sem{\decode(\memheap(\addr))}(\Phi) & &                                   
+  \Phi & \step \sem{\decode(\memheap(\addr))}(\Phi) & &                                   
                                                               \arraycolsep=0pt
@@
                                               \end{cases}\\
-  \sem{\move{\refreg{r_1}}{\rv}}(\Phi)            & = \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}}
+  \sem{\move{\refreg{r_1}}{\rv}}(\Phi)            & = 
+                                                    \begin{cases}
+                                                      \stdUpdatePc{\updateReg{r_1}{\rv}} & \rv \in \ints \\
+                                                      \stdUpdatePc{\updateReg{r_1}{\memreg(\rv)}} & \text{otherwise}
+                                                    \end{cases}
   \\
@@
                                                               \quad\text{and in either case $n_i \in \ints$}\\
-                                                              \quad\text{and $n_1 < n_2$}\\                                                              
-                                                            \end{array}\\
+                                                              \quad\text{and $n_1 < n_2$}\\                                                                                                            \end{array}\\
+                                                            \stdUpdatePc{\updateReg{r_1}{0}} &
+                                                            \arraycolsep=0pt
+                                                            \begin{array}[t]{l}
+                                                              \text{if for $i \in \{1,2\}$}\\
+                                                              \quad\text{$n_i = \rv_i$ or $n_i = \memreg(\rv_i)$}\\
+                                                              \quad\text{and in either case $n_i \in \ints$}\\
+                                                              \quad\text{and $n_1 \not< n_2$}\\                                                                                                            \end{array}\\
                                                             \failed & \text{otherwise}
@@
               \subsection{Macro definitions}
+In the following, we describe each of the macros.
+The descriptions are so detailed that it should be a simple matter to implement the macros.
+We provide a proposed implementation for each of the macros in order to install some confidence in the fact that it is possible to implement each of the macro.
               \begin{description}
@@
 By $\memSat[n_4]{\ms_3}{W_3}$ and $W_3 \futurestr W_2$, we know $\ms_\cls \subseteq \ms_3^p$, so using Hyp-Act(\ref{eq:g1-hyp-act}), we get $j$ such that
-                  \[
-                    \forall \ms_f \ldotp (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f) \step[j] (\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t \uplus \ms_f)
-                  \]
+
+                  \begin{multline}
+                    \forall \ms_f \ldotp (reg_3\update{\pcreg}{\updatePcPerm{c_\cls}},\ms_3^p \uplus \ms_3^t \uplus \ms_f) \step[j]\\ (\reg_3\update{\pcreg}{\updatePcPerm{c_\cls}}\update{r_\env}{c_\env},\ms_3^p \uplus \ms_3^t \uplus \ms_f)
+                  \end{multline}
                   Using Lemma~\ref{lem:anti-red-obs} it suffices to show
@@
 We also know that it is a global capability, so we can use it with private future worlds.
-We have $W_5 = \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset \uplus \ms_\act),\iota^\pwl(\dom(\ms_\unused')),\iota^\sta(\temp,\ms_f)] \futurestr W_3$, for some $\ms_\act$ and $\ms_\unused'$.
+We have $W_5 = \revokeTemp{W_4}[\iota^\sta (\temp,\emptyset \uplus \ms_\act \uplus \ms_r),\iota^\pwl(\dom(\ms_\unused'))] \futurestr W_3$, for some $\ms_\act$ and $\ms_\unused'$.
 By the execute/enter condition, we have
@@
               \[
-                \forall r \in \dom(W_3) \ldotp W_3 = W_{10}
+                \forall r \in \dom(W_3) \ldotp W_3(r) = W_{10}(r)
               \]
@@
               In other words, all the regions in $W_3$ have only taken public transitions compared to the corresponding regions in $W_{10}$.
+
+              The relation between the relevant worlds is sketched out in Figure~\ref{fig:worlds-in-awk-ex}.
 
@@
   \end{enumproof}
-
+\begin{figure}[t]
+  \centering
+\begin{tikzpicture}[main node/.style={}, node distance=1.5cm]
+  \node[main node] (1) {$W_3$};
+  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f4} called}}};
+  \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};
+
+  \node[main node] (2) [above right of=1]{$W_4$};
+  \node[main node] (label2) [above of=2,yshift=-1cm] {\textit{\footnotesize{first callback}}};
+
+  \node[main node,left] (const) [above of=given,yshift=-0.4cm] {\parbox{1.5cm}{\footnotesize{Constructed:}}};
+
+  \node[main node] (3) [below right of=2]{$W_6$};
+  \node[main node] (label3) [below of=3,yshift=0.75cm] {\footnotesize{\textit{callback returns}}};
+
+  \node[main node] (4) [above right of=3]{$W_7$};
+  \node[main node] (label4) [above of=4,yshift=-0.75cm] {\footnotesize{\textit{second callback}}};
+
+  \node[main node] (5) [below right of=4]{$W_8$};
+  \node[main node] (label5) [below of=5,yshift=1cm] {\footnotesize{\textit{callback returns}}};
+
+  \node[main node] (6) [above right of=5]{$W_{10}$};
+  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f4} returns}}};
+  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
+       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
+       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
+       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
+       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
+\end{tikzpicture}
+  \label{fig:worlds-in-awk-ex}
+\end{figure}
 \end{proof}
@@
  \[
-   W' \futurestr W \Leftrightarrow \xi(W') \futurestr \xi(W')
+   W' \futurestr W \Leftrightarrow \xi(W') \futurestr \xi(W)
  \]

### 2017-11-23 fully-abstract-comp/technical-report.tex
added a bit about the public transitions
Fix typo.
comment inserted
Resolved Lars' comment wrt K-relation
Added monotonicity requirements for memory interpretation functions in regions.
Proposed a definition for P in the LR
removed comments that we dealt with

@@
 \newcommand{\parfun}{\rightharpoonup}
+\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
 
@@
 \newcommand{\pubft}{\future^{\pub}}
+\newcommand{\monprivnefun}{\xrightarrow[\text{\tiny{$\privft$}}]{\textit{\tiny{mon, ne}}}}
+\newcommand{\monpubnefun}{\xrightarrow[\text{\tiny{$\pubft$}}]{\textit{\tiny{mon, ne}}}}
 
@@
 \[
+  \pi_\baddr(\vsc) =
+  \begin{cases}
+    \baddr & \text{if $\vsc = ((\_,\_),\baddr,\_,\_)$} \\ 
+    \pi_\baddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \stkptr{\_,\baddr,\_,\_}$}} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \retptrd{\_,(\baddr,\_)}$}} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \retptrc{\_,(\_,\_,\baddr,\_,\_)}$}}
+  \end{cases}
+\]
+\[
+  \pi_\eaddr(\vsc) =
+  \begin{cases}
+    \eaddr & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
+    \pi_\eaddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrd{\_,(\_,\eaddr)}$}} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrc{\_,(\_,\_,\_,\eaddr,\_)}$}}
+  \end{cases}
+\]
+\[
+  \pi_\lin(\vsc) = 
+  \begin{cases}
+    \lin & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
+    \pi_\lin(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
+    \src{\linear} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
+    \src{\linear} & \src{\text{if $\vsc = \retptrd{\_,(\_,\_)}$}} \\ 
+    \src{\lin} & \src{\text{if $\vsc = \retptrc{\_,(\_,\lin,\_,\_,\_)}$}}
+  \end{cases}
+\]
+
+\[
   \updPcAddr{c} \defeq \dots
@@
 \subsection{Worlds}
-\lau{30-10-2017: TODO: solve recursiveness in domain equation in the following. }
-\lau{02-11-2017: We may need revoked regions again to handle the stack.}
-\lau{06-11-2017: If we need to introduce revoked regions like we had for the local capabilities, we may consider the following: For the future world relation require the existence of a bijection from region names to region names.
-For all regions in the present world the future region is given by the bijection.
-This would mean that reinstantiation would happen simply by giving a bijection that maps to the new region.
-In other words, the weird shadowing would not be necessary.
-  \begin{mathpar}
-    \inferrule{ \dom(W') \supseteq \dom(W) \\ \exists m : \RegionName \fun \RegionName, \text{ $m$ injective} \ldotp \forall r \in \dom(W) \ldotp W'(m(r)) \pubft W(r) }
-    { W' \pubft W }
-  \end{mathpar}
-}
-\lau{10-11-2017: The above may not work well here because we do not want a $\spatialo$ region to be taken over by a $\pure$ region (as this would allow for normal capabilities addressing an area of memory that was only addressable by linear capabilities before).}
 \[
@@
   \begin{array}{l}
-    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2}) \uplus \\
-    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2})\uplus \\
-    \{\pure \} \times \State \times \Rels \times (\State \fun \World \fun \UPred{\MemSeg^2}) \uplus\\
+    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2})\uplus \\
+    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2}) \uplus\\
     \{\revoked\}
@@
 \lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
-\lau{26-20-2017: It is not clear to me what the monotonicity requirements on the memory interpretation functions should be.}
   \[
@@
 \end{mathpar}
-
+%
 The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
@@
             { (v,s',\phi^\pub,\phi,H) \privft (v,s,\phi^\pub,\phi,H)  }
-
+\end{mathpar}
+%
+Public future regions are allowed to transition according to the public part of their transition system.
+\begin{mathpar}
+    \inferrule{ }
+            { \revoked \pubft \revoked }
+\and
+  \inferrule{ (s,s') \in \phi^\pub }
+            { (v,s',\phi^\pub,\phi,H) \pubft (v,s,\phi^\pub,\phi,H)  }
 \end{mathpar}
@@
       \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair[.]{k_{d,S},k_{c,S}}{k_{d,T},k_{c,T}}} \in \lrk(W_\lrk) \wedge \\
+      \qquad\npair[n']{\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}} \in \lrp(W) \wedge \\
+      \qquad\npair[n']{\stpair[.]{k_{c,S},k_{d,S}}{k_{c,T},k_{d,T}}} \in \lrk(W_\lrk) \wedge \\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
@@
 \lau{03-11-2017: TODO take a look at linearity of this definition - in particular the recovered pc and stack pointer.}
-\[
-  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\k_{d,T},k_{c,T}\endarray\right)} \middle|
+
+\[
+  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\k_{c,T},k_{d,T}\endarray\right)} \middle|
     \begin{array}{l}
@@
       c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
+      \npair[n]{\stpair[.]{\retptrc(\sigma,\_)}{k_{c,T}},\stpair[.]{\retptrc{\retptrd(\sigma,\_)}}{k_{d,T}}} \in \lrp(W) \wedge \\
       \forall n' \leq n, W' \pubft W \ldotp \\
@@
       \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
-      \qquad\npair[n']{\src{\retptrd(\sigma,\_)},\src{\retptrc(\sigma,\_)},k_{d,T},k_{c,T}} \in \lrp(W') \wedge \\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
@@
   \lrp(W) = \left\{
-    \npair{\stpair{v}{v}} \middle| \text{make sure linearity constraint satisfied}\dots 
+    \npair{\stpair[.]{c_1}{c_1'}\stpair[.]{c_2}{c_2'}} \middle| 
+    \begin{array}{l}
+      \text{for $i \in \{1,2\}$} \\
+      \src{\baddr_i} = \pi_\baddr(\src{c_i}) \wedge \src{\eaddr_i} = \pi_\eaddr(\src{c_i}) \wedge 
+      \baddr_i' = \pi_\baddr(c_i') \wedge \eaddr_i' = \pi_\eaddr(c_i') \wedge \\
+      \src{\lin_i} = \pi_\lin(\src{c_i}) = \pi_\lin(c_i') \\
+      \quad \exists r \in \addressable{l,W} \ldotp\\
+      \qquad \forall \npair{n}{\stpair{\ms}{\ms}} \in W(r).H \; W(t).s \xi^{-1}(W)  \ldotp \\
+      \qquad \quad \dom(\src{\ms_S}) \subseteq [\src{\baddr_i},\src{\eaddr_i}] \wedge \dom(\ms_T) \subseteq [\baddr_i',\eaddr_i']      
+    \end{array}
   \right\}
@@
       \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
-      \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{(c_{d,S},c_{d,T})} \in \lrv(W'') \wedge \\
+      \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
       \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lrp(W' \oplus W'') \Rightarrow \\
-      \quad \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lre(W' \oplus W'')
+      \quad \qquad \npair[n']{\stpair[.]{((\perm,lin),\baddr,\eaddr,\aaddr)}{((\perm,lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
     \end{array}

### 2017-11-23 fully-abstract-comp/technical-report.tex
more small changes: removed comment, added text, and rearranged some things

@@
 \section{Logical Relation}
+In the following definitions, \src{blue} is used to indicate values related to the source machine.
+This is unlike previous definitions, where \src{blue} was used to indicate source language specific parts of definitions.
 \subsection{Worlds}
@@
 $\spatial$ signifies that the region is not owned and can thus not be addressed.
+At the same time it signifies that if something else addresses it, it is a $\linear$ capability.
 Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities.
@@
   \[
-   \Rels = \{\phi_\pub \times \phi \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
+   \Rels = \{(\phi_\pub, \phi) \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
   \]
@@
   (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
+  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
   (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
-                                           =  & \; (\spatialo,\var{sm}) \\
-  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
+                                           =  & \; (\spatialo,\var{sm})
 \end{align*}
@@
 Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region.
-It is further not defined if the two sides do not agree on region type and state machine.
+It is further not defined if the two sides do not agree on region type or state machine.
 Here $\var{sm}$ is the state machine of the region.
@@
   \inferrule{ (s,s') \in \phi }
-            { (v,s',\phi^\pub,\phi,H) \privft (v,s,\phi^\pub,\phi,H)  }
+            { (v,s',\phi_\pub,\phi,H) \privft (v,s,\phi_\pub,\phi,H)  }
 \end{mathpar}
@@
 \and
-  \inferrule{ (s,s') \in \phi^\pub }
-            { (v,s',\phi^\pub,\phi,H) \pubft (v,s,\phi^\pub,\phi,H)  }
+  \inferrule{ (s,s') \in \phi_\pub }
+            { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
 \end{mathpar}
@@
 \]
-In the definition of memory satisfaction, $R_\ms$ acts as a partitioning in the two related memories.
+In the definition of memory satisfaction, $R_\ms$ creates a partitioning of the two related memories.
 As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language.
@@
 \lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
-\lau{19-10-2017: Maybe linearity should be taken care of in just one place (instead of in $\lrr$ relation and mem satisfaction).
-That is, add a predicate that checks whether the configuration satisfies linearity (such a condition may be to syntactic to work?).}
-In the following definitions, \src{blue} is used to indicate values related to the source machine.
-This is unlike previous definitions, where \src{blue} is used  indicate source language specific parts of definitions.
 \[

### 2017-11-24 fully-abstract-comp/technical-report.tex
Removed two lines from the memory satisfaction definition. We will reintroduce them if we need them.
Added a frame to the observation relation.
Updated the K, R (and P) relation
P had a couple of typos.
K now assumes fewer things.
R is now indexed with a set R of registers that are not taken into consideration as K and E has different needs.
Typo in memSat, changed P-LR to take list, specified what we mean by R being a separation of W in R-relation.
wrote up some suggestions for updates of V-relation based on chat with Lars

@@
 \newcommand{\nsupeq}[1][n]{\overset{#1}{\supseteq}}
+\newcommand{\nequal}[1][n]{\overset{#1}{=}}
 
@@
 \newcommand{\monpubnefun}{\xrightarrow[\text{\tiny{$\pubft$}}]{\textit{\tiny{mon, ne}}}}
+
+%%% Regions
+\newcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
+\newcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
+\newcommand{\spa}{\mathrm{s}}
+\newcommand{\spao}{\mathrm{so}}
+\newcommand{\pur}{\mathrm{p}}
 
@@
       \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
-      \quad \exists r_0 , \dots, r_m, r_{m+1} \ldotp \pi_1(R_\ms(r_0,\dots,r_m)) = \ms_{\stk,0},\dots , \ms_{\stk,m} \wedge \\
-      \qquad \pi_1(R_\ms(r_{m+1})) = \ms_\stk \wedge \\
-      \qquad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
-      \qquad \exists R_w : \dom(W) \fun \World \ldotp\\
-      \quad \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
-      \quad \qquad \forall r \in \dom(W) , n' < n \ldotp \\
-      \qquad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
+      \quad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \uplus \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_w : \dom(W) \fun \World \ldotp\\
+      \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
+      \qquad \forall r \in \dom(W) , n' < n \ldotp \\
+      \quad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
     \end{array}
@@
       \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}} \in \lrp(W) \wedge \\
+      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
       \qquad\npair[n']{\stpair[.]{k_{c,S},k_{d,S}}{k_{c,T},k_{d,T}}} \in \lrk(W_\lrk) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rretd,\rretc,\rdata\})(W_\lrr) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
     \begin{array}{l}
-      \forall i \leq n \ldotp \\
-      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T,\reg_T) \term\\
+      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
+      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
       \quad \tand\\
-      \quad (\ms_T,\reg_T) \term[i] \Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \term
+      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term
     \end{array}
@@
 \]
-
-\lau{03-11-2017: TODO take a look at linearity of this definition - in particular the recovered pc and stack pointer.}
 
@@
       c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
-      \npair[n]{\stpair[.]{\retptrc(\sigma,\_)}{k_{c,T}},\stpair[.]{\retptrc{\retptrd(\sigma,\_)}}{k_{d,T}}} \in \lrp(W) \wedge \\
+      \npair[n]{[\stpair[.]{\retptrc(\sigma,\_)}{k_{c,T}},\stpair[.]{\retptrd(\sigma,\_)}{k_{d,T}}]} \in \lrp(W) \wedge \\
       \forall n' \leq n, W' \pubft W \ldotp \\
       \quad \forall W_\lrr, W_\lrm \ldotp  \\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
       \qquad\opc = c_c \wedge\\
       \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
-      \qquad\src{reg_S}(\rstk) = \src{\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\_}} \wedge \\
-      \qquad\eaddr_\stk = \aaddr_d-1\\
+      \qquad\npair[n']{\array{l}\src{\reg_S\update{\rstk}{\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}}},\\
+                               {\reg_T\update{\rstk}{((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)}}\endarray}\in \lrr(\{r_\stk\})(W_\lrr) \wedge\\ 
+      \qquad\npair[n']{[(\opc,\opc)]} \in \lrp(W')\\
       \qquad\Rightarrow\\
-      \qquad\npair[n']{\left(\array{l}\src{\left(\arraycolsep=0pt\array{l}\ms_S,\reg_S\updReg{\pcreg,\rstk}{\opc,\stkptr{\rw,\baddr_\stk,\eaddr_d,\aaddr_d}},\\\stk, \ms_\stk \uplus \ms_{\stk,f}\endarray\right)},\\
-                                          (\ms_T,\reg_T\updReg{\pcreg,\rstk}{\opc,((\rw,\lin),\baddr_\stk,\eaddr_d,\aaddr_d)})\endarray\right)}
-      \in \lro(W' \oplus W_\lrr \oplus W_\lrm)
+      \qquad \npair[n']{\left(\arraycolsep=0pt\array{l}
+                          \left(\arraycolsep=0pt\array{l}
+                            \src{\ms_S,\reg_S\update{\pcreg,\rstk}{\opc,\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}},}\\
+                            \src{\stk, \ms_\stk \uplus \ms_{\stk,f}}
+                          \endarray \right) \\
+                          \left(\ms_T,\reg_T\update{\pcreg,\rstk}{\opc,((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)} \right)
+                        \endarray \right)} \in \lro(W'\oplus W_\lrr \oplus W_\lrm)
     \end{array}
@@
   \lrp(W) = \left\{
-    \npair{\stpair[.]{c_1}{c_1'}\stpair[.]{c_2}{c_2'}} \middle| 
+    \npair{ \overline{\stpair[.]{c}{c'}} } \middle| 
     \begin{array}{l}
-      \text{for $i \in \{1,2\}$} \\
-      \src{\baddr_i} = \pi_\baddr(\src{c_i}) \wedge \src{\eaddr_i} = \pi_\eaddr(\src{c_i}) \wedge 
-      \baddr_i' = \pi_\baddr(c_i') \wedge \eaddr_i' = \pi_\eaddr(c_i') \wedge \\
-      \src{\lin_i} = \pi_\lin(\src{c_i}) = \pi_\lin(c_i') \\
+      \text{for $\stpair[.]{c}{c'} \in \overline{\stpair[.]{c}{c'}}$} \\
+      \src{\baddr} = \pi_\baddr(\src{c}) \wedge \src{\eaddr} = \pi_\eaddr(\src{c}) \wedge 
+      \baddr = \pi_\baddr(c) \wedge \eaddr = \pi_\eaddr(c) \wedge \\
+      \src{\lin} = \pi_\lin(\src{c}) = \pi_\lin(c) \\
       \quad \exists r \in \addressable{l,W} \ldotp\\
-      \qquad \forall \npair{n}{\stpair{\ms}{\ms}} \in W(r).H \; W(t).s \xi^{-1}(W)  \ldotp \\
-      \qquad \quad \dom(\src{\ms_S}) \subseteq [\src{\baddr_i},\src{\eaddr_i}] \wedge \dom(\ms_T) \subseteq [\baddr_i',\eaddr_i']      
+      \qquad \forall \npair{\stpair{\ms}{\ms}} \in W(r).H \; W(r).s \; \xi^{-1}(W)  \ldotp \\
+      \qquad \quad \dom(\src{\ms_S}) \subseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \subseteq [\baddr,\eaddr]      
     \end{array}
@@
 \[
-  \lrr(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+  \lrr(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
     \begin{array}{l}
       \exists S : \RegName \fun \World \ldotp \\
-      \quad S \text{ is a separation of $W$} \tand\\
-      \quad \forall r \in \RegName \setminus \{\pcreg, \rretd,\rretc,\rdata \}\ldotp\\
+      \quad W = \bigoplus_{r \in \RegName} S(r) \\
+      \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
       \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
@@
 Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
-
-\[
-  \lrv(W) =
+\lau{TODO: Rewrite the other definitions to pass around $H_\sigma$ (if this is how we want to handle seals.)}
+\[
+  \lrv(H_\sigma,W) =
   \begin{array}[t]{l}
     \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+%
     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\_)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
     \begin{array}{l}
-    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W_\lrv) \ldotp\\
-      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
+%      \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W_\lrv) \ldotp\\
+%      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
     \end{array}
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\_)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \;\middle| \;
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \;\middle| \;
     \begin{array}{l}
-    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W_\lrv) \ldotp\\
-      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
+      \vsc_D = ((\rw,\linear),\baddr,\eaddr,\baddr-1) \wedge\\
+      \exists r \in \addressable{\linear,W} \ldotp \\
+      \quad  \exists \stpair{\ms}{\ms} \ldotp \\
+      \qquad \dom(\ms) = [\baddr, \eaddr] \wedge \stareg{\spao} \nequal W(r) \wedge \\
+      \qquad \forall W' \privft W \ldotp \\
+      \qquad \quad \exists r' \ldotp W'(r') = W(r) \wedge \exists r'' \ldotp W'(r'') = \stdreg{[\stkb,\aaddr]}{\spao} \ldotp \\
+      \qquad \qquad { \color{red} \forall (c, c') \in H_\sigma (\sigma) \ldotp \npair{\src{c,\retptrd(\sigma,\baddr,\eaddr)}, c' , \sealed{\sigma,\vsc_D}} \in \lrk(W') }%\text{\lau{This needs some work - specifically what i c and c' and do we need to seal them?}}
+
+%    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W_\lrv) \ldotp\\
+%      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
     \end{array}
     \right\}\cup \\
+%
     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
     \begin{array}{l}
-
+      \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
+      \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
+      \quad \npair{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lre(W) \wedge \\
+      \quad \npair{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lre(W)
     \end{array}
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} \;\middle|\;
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
     \right\} \cup \\
+%
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \begin{array}{r l l}
+      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
+      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
+    \end{array}
     \right\} \cup \\
+%
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
@@
       \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
-      \qquad \npair[n']{((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,S},((\perm,lin),\baddr,\eaddr,\aaddr),c_{d,T}} \in \lrp(W' \oplus W'') \Rightarrow \\
+      \qquad \npair[n']{[\stpair[.]{((\perm,lin),\baddr,\eaddr,\aaddr)}{((\perm,lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}]} \in \lrp(W' \oplus W'') \Rightarrow \\
       \quad \qquad \npair[n']{\stpair[.]{((\perm,lin),\baddr,\eaddr,\aaddr)}{((\perm,lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')

### 2017-11-28 presentations/presentation-illustrations.tex
Start work on some tikz figures for making presentations.

\documentclass{beamer}

\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{positioning}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.pathreplacing}
\usepackage{tikzpeople}

% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
\makeatletter
\newif\if@showgrid@grid
\newif\if@showgrid@left
\newif\if@showgrid@right
\newif\if@showgrid@below
\newif\if@showgrid@above
\tikzset{%
    every show grid/.style={},
    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
    show grid/.default={true},
    show grid/.cd,
    labels/.style={font={\sffamily\small},help lines},
    xlabels/.style={},
    ylabels/.style={},
    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
    true/.style={left,below},
    false/.style={left=false,right=false,above=false,below=false,grid=false},
    none/.style={left=false,right=false,above=false,below=false},
    all/.style={left=true,right=true,above=true,below=true},
    grid/.is if=@showgrid@grid,
    left/.is if=@showgrid@left,
    right/.is if=@showgrid@right,
    below/.is if=@showgrid@below,
    above/.is if=@showgrid@above,
    false,
}

\def\@showgrid#1{%
    \begin{scope}[every show grid,show grid/.cd,#1]
    \if@showgrid@grid
    \begin{pgfonlayer}{background}
    \draw [help lines]
        (current bounding box.south west) grid
        (current bounding box.north east);
%
    \pgfpointxy{1}{1}%
    \edef\xs{\the\pgf@x}%
    \edef\ys{\the\pgf@y}%
    \pgfpointanchor{current bounding box}{south west}
    \edef\xa{\the\pgf@x}%
    \edef\ya{\the\pgf@y}%
    \pgfpointanchor{current bounding box}{north east}
    \edef\xb{\the\pgf@x}%
    \edef\yb{\the\pgf@y}%
    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
    \if@showgrid@below
    \foreach \X in {\xbeg,...,\xend} {
        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
    }
    \fi
    \if@showgrid@above
    \foreach \X in {\xbeg,...,\xend} {
        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
    }
    \fi
    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
    \if@showgrid@left
    \foreach \Y in {\ybeg,...,\yend} {
        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
    }
    \fi
    \if@showgrid@right
    \foreach \Y in {\ybeg,...,\yend} {
        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
    }
    \fi
    \end{pgfonlayer}
    \fi
    \end{scope}
}
\makeatother

\begin{document}
\begin{frame}
  \frametitle{Traditional stack pointers}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
    \end{scope}
    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);

    % traditional:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);

    % traditional call:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);

    % traditional return:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);

    % traditional call, attack:
    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };

    % read other stack frames
    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
    \draw<.> (3,7) node[devil,mirrored,minimum size=1cm] {};
    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);

    % break well-bracketedness
    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
      skip caller frame};
  \end{tikzpicture}

\end{frame}

\begin{frame}
  \frametitle{Stack and return capabilities}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
    \end{scope}

    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);

    % first step: capabilities
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15);
      \draw<.>[fill=white] (0,5.5) rectangle (6,15.1);
    \end{scope}
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};

    % call using capabilities
    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15);
      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
    \end{scope}
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);

    % attacker...
    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15);
      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
    \end{scope}
    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };

    \begin{scope}
      \clip (13.9,6) rectangle (18.1,13);
      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
    \end{scope}
    \draw<.> (16,14) node {heap memory};
    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);

    \begin{scope}
      \clip (13.9,6) rectangle (18.1,13);
      \draw<+>[fill=gray!20] (14,5.9) rectangle (18,13.1);
    \end{scope}
    \draw<.> (16,14) node {heap memory};
    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (10.5,6.5) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};

    \begin{scope}
      \clip (13.9,6) rectangle (18.1,13);
      \draw<+>[fill=gray!20] (14,5.9) rectangle (18,13.1);
    \end{scope}
    \draw<.> (16,14) node {heap memory};
    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (10.5,6.5) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
    \draw<.>[fill=gray!50] (0,2.5) rectangle (6,9.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.> (3,12) node[devil,mirrored,minimum size=1cm] {};
    \draw<.> (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15);
      \draw<.>[fill=white] (0,13.5) rectangle (6,15.1);
    \end{scope}
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw<.> node[right=0cm of sp1] { stack pointer };
    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \draw<.>[fill=gray!50] (0,9.5) rectangle node {\scriptsize old return pointer data} (6,10);
    \draw<.>[fill=gray!50] (0,10) rectangle node {\scriptsize old return pointer code} (6,10.5);
    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);

  \end{tikzpicture}
\end{frame}
\end{document}

### 2017-11-28 fully-abstract-comp/technical-report.tex
added the retptr code case and fixed some typos. Added the standard static region definition. More typos.

@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\_)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
     \begin{array}{l}
-%      \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrd(\sigma,\_)}{\sealed{\sigma,\vsc_D}}} \in \lrv(W_\lrv) \ldotp\\
-%      \npair[n']{\retptrc(\sigma,\_),\retptrd(\sigma,\_),\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
+      \vsc_c = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge \\
+      \forall (\src{c},c') \in H_\sigma(\sigma) \ldotp \\
+      \quad \exists r \in \addressable(\linear,W) \ldotp \\
+      \qquad  \exists \stpair{\ms}{\ms} \ldotp \\
+      \qquad \quad \dom(\src{\ms_S}) = \dom(\ms_T) = [\pi_\baddr(c), \pi_\eaddr (c)] \wedge \stareg{\spao} \nequal W(r) \wedge \\
+      \qquad \qquad \forall W' \privft W \ldotp \\
+      \qquad \qquad \quad \exists r', W'(r') = W(r), \exists r'', W'(r'') = \stdreg{[\stkb,\eaddr]}{\spao} \ldotp \\
+      \qquad \qquad \qquad { \color{red} \npair{(\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)}, \src{c}, \sealed{\sigma,\vsc_c}, c')}} \in \lrk(W')
     \end{array}
@@
       \quad  \exists \stpair{\ms}{\ms} \ldotp \\
-      \qquad \dom(\ms) = [\baddr, \eaddr] \wedge \stareg{\spao} \nequal W(r) \wedge \\
+      \qquad \dom(\src{\ms_S}) = \dom(\ms_T) = [\baddr, \eaddr] \wedge \stareg{\spao} \nequal W(r) \wedge \\
       \qquad \forall W' \privft W \ldotp \\
-      \qquad \quad \exists r' \ldotp W'(r') = W(r) \wedge \exists r'' \ldotp W'(r'') = \stdreg{[\stkb,\aaddr]}{\spao} \ldotp \\
-      \qquad \qquad { \color{red} \forall (c, c') \in H_\sigma (\sigma) \ldotp \npair{\src{c,\retptrd(\sigma,\baddr,\eaddr)}, c' , \sealed{\sigma,\vsc_D}} \in \lrk(W') }%\text{\lau{This needs some work - specifically what i c and c' and do we need to seal them?}}
+      \qquad \quad \exists r', W'(r') = W(r), \exists r'', W'(r'') = \stdreg{[\stkb,\eaddr]}{\spao} \ldotp \\
+      \qquad \qquad { \color{red} \forall (\src{c}, c') \in H_\sigma (\sigma) \ldotp \npair{\src{c,\retptrd(\sigma,\baddr,\eaddr)}, c' , \sealed{\sigma,\vsc_D}} \in \lrk(W') }%\text{\lau{This needs some work - specifically what i c and c' and do we need to seal them?}}
 
@@
     \end{array}
-
     \right\}
@@
 \]
+\lau{28-11-2017: Do we want the execute condition for $\rwx$ is we are doing write xor execute?}
 
@@
   \begin{cases}
-    \{ r \mid W(r) = (\normal,\_) \} & \text{if $\lin = \normal$} \\
+    \{ r \mid W(r) = (\pure,\_) \} & \text{if $\lin = \normal$} \\
     \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad W(r) \nsubeq \iota_{[\baddr',\eaddr']}
+      \qquad W(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
     \end{array}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad W(r) \nsupeq[n-1] \iota_{[\baddr',\eaddr']} \wedge \\
+      \qquad W(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
       \qquad W(r) \text{ is address-stratified}
@@
 \[
-  \iota_A \defeq (\normal,*,=,=,H_A)
-\]
+  \stdreg{A}{v} \defeq (v,*,=,=,H_A)
+\]
+for readability, we use $\spao$ short for $\spatialo$, $\spa$ short for $\spatial$, and $\pur$ as short for $\pure$.
 \lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
@@
       \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
-      \forall \aaddr \in A \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(\xi(\hat{W}))
+      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \uplus_{\aaddr \in \dom(\ms_S)} S(\aaddr) \wedge\\
+      \quad \forall \aaddr \in A \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
 \lau{01-11-2017: TODO look into what kind of world separation should happen here.}
+
+\[
+  \stareg{v} = (v,*,=,=,H^\mathrm{sta}_{\stpair{\ms}{\ms}})
+\]
+\[
+  H^\mathrm{sta}_{\stpair{\ms}{\ms}} = \{ \npair{\stpair{\ms}{\ms}} \}
+\]
 

### 2017-11-28 presentations/presentation-illustrations.tex
More illustrations.

@@
     % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
     \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
     \begin{scope}
@@
 
+    % stores stack pointer in heap
     \begin{scope}
@@
 
+    % attacker returns
     \begin{scope}
@@
     \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
+    \draw<.>[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
 
+    % attacker gets called again + uses old stack pointer.
     \begin{scope}
@@
     \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
-    \draw<.>[fill=gray!50] (0,2.5) rectangle (6,9.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
     \draw<.>[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
@@
     \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.>[fill=gray!50] (0,9.5) rectangle node {\scriptsize old return pointer data} (6,10);
-    \draw<.>[fill=gray!50] (0,10) rectangle node {\scriptsize old return pointer code} (6,10.5);
     \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
@@
     \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-
   \end{tikzpicture}
 \end{frame}
+
+\begin{frame}
+  \frametitle{Stack and return capabilities: another attack}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % first step: capabilities
+    \draw<+>[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15);
+      \draw<.>[fill=white] (0,4) rectangle (6,15.1);
+    \end{scope}
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw<.> node[right=0cm of sp1] { stack pointer };
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+
+    % attacker stores copy of stack pointer high in the stack
+    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15);
+      \draw<.>[fill=white] (0,7) rectangle (6,15.1);
+    \end{scope}
+    \draw<.> (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw<.> node[right=0cm of sp1] { stack pointer };
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
+    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
+    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
+
+    % attacker calls trusted code again
+    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \draw<.>[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw<.> node[right=0cm of sp1] { stack pointer };
+    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);
+
+    % attacker calls trusted code again
+    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \draw<.>[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \draw<.>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
+    \draw<.> (3,10.5) node[devil,mirrored,minimum size=1cm] {};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw<.> node[right=0cm of sp1] { stack pointer };
+    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);
+    \draw<.> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack Capabilities Problems}
+  \begin{itemize}
+  \item Stack capabilities not a panacea
+  \end{itemize}
+\end{frame}
 \end{document}

### 2017-11-28 presentations/presentation-illustrations.tex

@@
 \usepackage{pgf}
+\usepackage{ulem}
 \usepackage{tikz}
@@
 \usetikzlibrary{decorations.pathreplacing}
+\usetikzlibrary{arrows,shapes}
 \usepackage{tikzpeople}
@@
 
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack and return capabilities: Attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
     % attacker...
@@
 \begin{frame}
-  \frametitle{Stack and return capabilities: another attack}
+  \frametitle{Stack and return capabilities: Attack 2}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \draw<.>[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15);
+      \draw<.>[fill=white] (0,9) rectangle (6,15.1);
+    \end{scope}
     \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \draw<.>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15);
+      \draw<.>[fill=white] (0,12) rectangle (6,15.1);
+    \end{scope}
     \draw<.> (3,10.5) node[devil,mirrored,minimum size=1cm] {};
@@
   \begin{itemize}
-  \item Stack capabilities not a panacea
+  \item Stack capabilities are not a panacea
+  \item Stack pointers must not leave stack frame (through memory or stack)
+  \item Similar problems for return pointers (omitted)
+  \item Solution: local capabilities?
+    \begin{itemize}
+    \item Make stack and return capabilities local
+    \item Cannot leave registers (except on stack) (\sout{Attack 1})
+    \item Stack clearing on boundary crossings (\sout{Attack 2})
+    \end{itemize}
   \end{itemize}
 \end{frame}
+
+\begin{frame}
+  \frametitle{Local Stack Capabilities prevent Attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % attacker...
+    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15);
+      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
+    \end{scope}
+    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw<.> node[right=0cm of sp1] { stack pointer };
+
+    % stores stack pointer in heap
+    \begin{scope}
+      \clip (13.9,6) rectangle (18.1,13);
+      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
+    \end{scope}
+    \draw<.> (16,14) node {heap memory};
+    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
+
+    \draw<.>[green,thick] (13,7.5) -- (19,9);
+    \draw<.>[green,very thick] (19,7.5) -- (13,9);
+    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
+
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack Clearing prevent Attack 2}
+  TODO
+\end{frame}
 \end{document}

### 2017-11-29 fully-abstract-comp/technical-report.tex
fixed a couple of typos
Added the non ret-token case in L lr. Not sure whether we will need this.
typos

@@
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
+    &\cbnf & \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
     &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
@@
       \cb = \seal{\linear,\_,\_,\_} \tor\\
-      \cb = \sealed{\cb',\_} \tand \isLinear{\cb'} 
+      \cb = \sealed{\_,\cb'} \tand \isLinear{\cb'} 
     \end{array}\\
@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_1) = \sealed{c_\code,\sigma} \tand \\
-                                       \Phi(r_2) = \sealed{c_\data,\sigma} \tand \\
+                                       \Phi(r_1) = \sealed{\sigma,c_\code} \tand \\
+                                       \Phi(r_2) = \sealed{\sigma,c_\data} \tand \\
                                        \nonExec{c_\data} \tand \\
@@
                                       \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
-                                      \vsc = \sealed{\Phi(r_2),\sigma} \tand\\
+                                      \vsc = \sealed{\sigma,\Phi(r_2)} \tand\\
                                       w_2 = \linCons{\Phi(r_2)}                                     
@@
                                                  \begin{array}[t]{l}
-                                                   \Phi(r_1) = \sealed{c_1,\sigma_1} \tand \\
-                                                   \Phi(r_2) = \sealed{c_2,\sigma_2} \tand \\
+                                                   \Phi(r_1) = \sealed{\sigma_1,c_1} \tand \\
+                                                   \Phi(r_2) = \sealed{\sigma_2,c_2} \tand \\
                                                    \sigma_1 = \sigma_2 \tand \\
@@
   \lau{27-09-2017: The ``somehow''s need to be fixed.}
-  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{c_\code,\sigma}$ and $\vsc_\data = \sealed{c_\data,\sigma}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
+  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{\sigma,c_\code}$ and $\vsc_\data = \sealed{\sigma,c_\data}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
   \begin{itemize}
@@
 \[
-  \lrk(W) = \left\{ \npair{\left(\array{l}\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\k_{c,T},k_{d,T}\endarray\right)} \middle|
+  \begin{array}{rl}
+  \lrk(W) = &\left\{ \npair{\left(\array{l}\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\k_{c,T},k_{d,T}\endarray\right)} \middle|
     \begin{array}{l}
@@
     \end{array}
+  \right\}\\ \cup
+    & \left\{ \npair{\left( k_{c,S},k_{d,S},k_{c,T},k_{d,T} \right)} \middle|
+      \begin{array}{l}
+        \pi_\sigma(k_{c,S}) = \pi_\sigma(k_{d,S}) \wedge \pi_\sigma(k_{c,T}) = \pi_\sigma(k_{d,T}) \wedge \npair{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \\
+        \forall n' \leq n, W' \pubft W \ldotp \\
+        \quad \npair[n']{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \Rightarrow\\
+        \qquad \forall \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W'_\lrr) \ldotp 
+               \forall \memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W'_\lrm} \ldotp\\
+        \qquad \quad \npair[n']{\left(
+        \arraycolsep=0pt\array{l}
+          (\src{\ms_S},\src{\reg_S}\update{\pcreg,\rdata}{\pi_\vsc(\src{k_{c,S}}),\pi_\vsc(\src{k_{d,S}})},\src{\stk},\src{\ms_\stk})\\
+          (\ms_T,\reg_T\update{\pcreg,\rdata}{\pi_\vsc(k_{c,T}),\pi_\vsc(k_{d,T})})
+        \endarray\right)} \in \lro(W' \oplus W'_\lrr \oplus W'_\lrm)
+      \end{array}
       \right\}
+  \end{array}
 \]
@@
 Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
+
 \lau{TODO: Rewrite the other definitions to pass around $H_\sigma$ (if this is how we want to handle seals.)}
@@
       \forall (\src{c},c') \in H_\sigma(\sigma) \ldotp \\
-      \quad \exists r \in \addressable(\linear,W) \ldotp \\
+      \quad \exists r \in \addressable{\linear,W} \ldotp \\
       \qquad  \exists \stpair{\ms}{\ms} \ldotp \\
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\lin,\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\lin,\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
     \right\} \cup \\
@@
       \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
-      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \uplus_{\aaddr \in \dom(\ms_S)} S(\aaddr) \wedge\\
+      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms_S)} S(\aaddr) \wedge\\
       \quad \forall \aaddr \in A \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))

### 2017-11-30 presentations/presentation-illustrations.tex
Refactor tikz code a bit.

--- 2017-11-28 presentations/presentation-illustrations.tex
@@
 
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack and return capabilities: Attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
     % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
+    \begin{onlyenv}<+-+(1)>
+      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
       \begin{scope}
         \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
+        \draw[fill=white] (0,9.5) rectangle (6,15.1);
       \end{scope}
-    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
 
+    % stores stack pointer in heap
+    \begin{onlyenv}<+>
       \begin{scope}
@@
       \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \end{onlyenv}
 
+    % attacker returns
     \begin{scope}
@@
     \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.>[fill=gray!50] (0,9.5) rectangle node {\scriptsize old return pointer data} (6,10);
-    \draw<.>[fill=gray!50] (0,10) rectangle node {\scriptsize old return pointer code} (6,10.5);
+    % \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    %   (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
+    % \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
     \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
@@
     \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack and return capabilities: Attack 2}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+      \draw[fill=white] (0,4) rectangle (6,16);
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % first step: capabilities
+    \begin{onlyenv}<+>
+      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+    \end{onlyenv}
+
+    % attacker stores copy of stack pointer high in the stack
+    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \begin{onlyenv}<.-.(2)>
+      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+      \draw<.-.(1)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
+      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
+      % \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      % (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+      % \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
+    \end{onlyenv}
+    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
+    \end{onlyenv}
+
+    % attacker calls trusted code again
+    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{onlyenv}<.>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % attacker uses old stack pointer
+    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \begin{onlyenv}<.-.(1)>
+      \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
+      \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
+    \end{onlyenv}
+    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack Capabilities Problems}
+  \begin{itemize}
+  \item Stack capabilities are not a panacea
+  \item Stack pointers must not leave stack frame (through memory or stack)
+  \item Similar problems for return pointers (omitted)
+  \item Solution: local capabilities?
+    \begin{itemize}
+    \item Make stack and return capabilities local
+    \item Cannot leave registers (except on stack) (\sout{Attack 1})
+    \item Stack clearing on boundary crossings (\sout{Attack 2})
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Local Stack Capabilities prevent Attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % attacker...
+    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15);
+      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
+    \end{scope}
+    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
+    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw<.> node[right=0cm of sp1] { stack pointer };
+
+    % stores stack pointer in heap
+    \begin{scope}
+      \clip (13.9,6) rectangle (18.1,13);
+      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
+    \end{scope}
+    \draw<.> (16,14) node {heap memory};
+    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
+
+    \draw<.>[green,thick] (13,7.5) -- (19,9);
+    \draw<.>[green,very thick] (19,7.5) -- (13,9);
+    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
 
@@
 \end{frame}
+
+\begin{frame}
+  \frametitle{Stack Clearing prevent Attack 2}
+  TODO
+\end{frame}
 \end{document}

### 2017-11-30 presentations/presentation-illustrations.tex
Some more tikz refactoring.

@@
       \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+      \begin{scope}
+        \clip (-.1,-.1) rectangle (6.1,15);
+        \draw[fill=white] (0,2.5) rectangle (6,15.1);
+      \end{scope}
     \end{scope}
@@
     % attacker...
-    \begin{onlyenv}<+-+(1)>
+    \begin{onlyenv}<+-+(2)>
       \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
       \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (6.1,15);
-        \draw[fill=white] (0,9.5) rectangle (6,15.1);
-      \end{scope}
       \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
@@
     % stores stack pointer in heap
-    \begin{onlyenv}<+>
+    \begin{onlyenv}<+->
       \begin{scope}
         \clip (13.9,6) rectangle (18.1,13);
-        \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
+        \draw[fill=gray!20] (14,5.9) rectangle (18,13.1);
       \end{scope}
-      \draw<.> (16,14) node {heap memory};
-      \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+      \draw (16,14) node {heap memory};
+      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+    \end{onlyenv}
     \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
     \end{onlyenv}
@@
     % attacker returns
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<+>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,6.5) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    \begin{onlyenv}<+>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.>[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
+    \end{onlyenv}
 
     % attacker gets called again + uses old stack pointer.
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<+>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,6.5) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
-    \draw<.>[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (3,12) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,13.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    \begin{onlyenv}<+>
+      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
+      \draw[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
+      \draw (3,12) node[devil,mirrored,minimum size=1cm] {};
+      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
         (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    % \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      \draw node[right=0cm of sp1] { stack pointer };
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    % \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
+      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
+    \end{onlyenv}
   \end{tikzpicture}
@@
       \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
-      \draw<.-.(1)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
         (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
       \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-      % \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      % (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      % \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
     \end{onlyenv}
@@
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
       \draw (6,13.25) edge[red,thick,->,in=45] (spold);
@@
       \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
       \draw node[right=0cm of sp1] { stack pointer };
-      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
     \end{onlyenv}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
Move illustrations to separate directory and add tikzpeople style files.

--- 2017-11-30 presentations/presentation-illustrations.tex
@@
     % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
+    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
     \begin{scope}
       \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
+      \draw[fill=white] (0,9.5) rectangle (6,15.1);
     \end{scope}
-    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
+    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
+    \draw node[right=0cm of sp1] { stack pointer };
 
@@
       \clip (13.9,6) rectangle (18.1,13);
-      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
+      \draw[fill=gray!20] (14,5.9) rectangle (18,13.1);
     \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \draw (16,14) node {heap memory};
+    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
 
-    \draw<.>[green,thick] (13,7.5) -- (19,9);
-    \draw<.>[green,very thick] (19,7.5) -- (13,9);
-    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
-
+    \draw[teal,thick] (13,7.5) -- (19,9);
+    \draw[teal,very thick] (19,7.5) -- (13,9);
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
   \end{tikzpicture}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
Slide about attack 2 being prevented.

@@
 \usetikzlibrary{backgrounds}
+\usetikzlibrary{calc}
 \usetikzlibrary{positioning}
@@
       \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+      \draw[fill=white] (0,9.5) rectangle (6,15.2);
     \end{scope}
@@
     \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
     \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
@@
 \begin{frame}
-  \frametitle{Stack Clearing prevent Attack 2}
-  TODO
+  \frametitle{Local caps: Stack clearing prevents Attack 2}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+      \draw[fill=white] (0,4) rectangle (6,16);
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % attacker stores copy of stack pointer high in the stack
+    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \begin{onlyenv}<+>
+      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+    \end{onlyenv}
+    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);
+
+    % attacker calls trusted code again
+    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{onlyenv}<.-.(1)>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % trusted code clears the stack
+    \begin{onlyenv}<+->
+      \foreach \x in {9,9.5,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
+    (13,13.5) -- (10,14.5);
+
+    % attacker uses old stack pointer
+    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \begin{onlyenv}<.-.(1)>
+      \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
+      \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+  \end{tikzpicture}
 \end{frame}
+
 \end{document}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
Fix spurious slide.
Add slides to show that trusted code needs to clear stack on return as well.
Improve explanation of call when using stack capabilities.

@@
     % first step: capabilities
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,5.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    \begin{onlyenv}<+-+(4)>
+      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(3)>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+    \end{onlyenv}
+
+    %prepare call: store old return pointer
+    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
+    \begin{onlyenv}<+-+(2)>
+      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
+      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
+      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
+      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { new stack pointer };
+    \end{onlyenv}
 
     % call using capabilities
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-
+    \begin{onlyenv}<+->
+      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
+    \end{onlyenv}
   \end{tikzpicture}
@@
     \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.-.(1)>
+    \begin{onlyenv}<.>
       \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
@@
     \end{onlyenv}
+
+    % attacker returns
+    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
+    \begin{onlyenv}<.>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % trusted code clears stack again
+    \begin{onlyenv}<+->
+      \foreach \x in {7,7.5,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+    % then, the trusted code can return
+    \begin{onlyenv}<+>
+      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+    \end{onlyenv}
+
   \end{tikzpicture}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
small changes to presentation illustration

@@
     % attacker...
-    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller satck frame};
     \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
@@
 
-    \draw[teal,thick] (13,7.5) -- (19,9);
-    \draw[teal,very thick] (19,7.5) -- (13,9);
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
+    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
+    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};
   \end{tikzpicture}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
Changed layout for adversary frames and added a macro for drawing adv frames.

@@
 \usepackage{tikzpeople}
+
+\newcommand{\drawadv}[3][]{
+      \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+      \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+      \draw[fill=white, fill opacity=0.8] #2 rectangle #3 node[pos=.5] {};
+      \draw #2 rectangle #3 node[pos=.5] {\footnotesize #1};
+}
+
+\newcommand{\drawinactiveadv}[3][]{
+      \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+      \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+      \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+      \draw #2 rectangle #3 node[pos=.5] {\footnotesize #1};
+}
 
@@
       \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee stack frame]{(0,6.5)}{(6,9.5)}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
       \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,12) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee stack frame]{(0,10.5)}{(6,13.5)}
       \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
@@
     \begin{onlyenv}<.-.(2)>
-      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee stack frame]{(0,4)}{(6,7)}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \begin{onlyenv}<.-.(1)>
-      \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
-      \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee (3) stack frame]{(0,9)}{(6,12)}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller satck frame};
-    \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
+    \drawadv[callee stack frame]{(0,6.5)}{(6,9.5)}
     \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \begin{onlyenv}<+>
-      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee stack frame]{(0,4)}{(6,7)}
     \end{onlyenv}
@@
     \begin{onlyenv}<.>
-      \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
-      \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
+      % \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
+      % \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee (3) stack frame]{(0,9)}{(6,12)}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \begin{onlyenv}<+>
-      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+      % \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+      % \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
+      \drawadv[callee stack frame]{(0,4)}{(6,7)}
     \end{onlyenv}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
fixed an issue with macro and added a couple more that can be used

@@
 
+\usepackage{ifthen}
 \usepackage{pgf}
@@
 
-\newcommand{\drawadv}[3][]{
+\newcommand{\actadv}[4][]{
+  \ifthenelse{\equal{#1}{}}{
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
     \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-      \draw[fill=white, fill opacity=0.8] #2 rectangle #3 node[pos=.5] {};
+    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1> ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw<#1>[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+
+  }
+}
+
+\newcommand{\inactadv}[3][]{
+  \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+  \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+  \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
   \draw #2 rectangle #3 node[pos=.5] {\footnotesize #1};
@@
 
-\newcommand{\drawinactiveadv}[3][]{
+\newcommand{\actsf}[3][]{
   \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-      \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+}
+
+\newcommand{\inactsf}[3][]{
+  \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
   \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
@@
     \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \draw<.> (3,7) node[devil,mirrored,minimum size=1cm] {};
+    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
     \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
@@
       \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \drawadv[callee stack frame]{(0,6.5)}{(6,9.5)}
+      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
       \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \drawadv[callee stack frame]{(0,10.5)}{(6,13.5)}
+      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
       \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
@@
     \begin{onlyenv}<.-.(2)>
-      \drawadv[callee stack frame]{(0,4)}{(6,7)}
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \begin{onlyenv}<.-.(1)>
-      \drawadv[callee (3) stack frame]{(0,9)}{(6,12)}
+      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller satck frame};
-    \drawadv[callee stack frame]{(0,6.5)}{(6,9.5)}
+    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
     \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \begin{onlyenv}<+>
-      \drawadv[callee stack frame]{(0,4)}{(6,7)}
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
     \end{onlyenv}
@@
     \begin{onlyenv}<.>
-      % \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
-      % \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
-      \drawadv[callee (3) stack frame]{(0,9)}{(6,12)}
+      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
     \begin{onlyenv}<+>
-      % \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-      % \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
-      \drawadv[callee stack frame]{(0,4)}{(6,7)}
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
     \end{onlyenv}

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
used the inactive adv command in the presentation

@@
 
-\newcommand{\inactadv}[3][]{
+\newcommand{\inactadv}[4][]{
+  \ifthenelse{\equal{#1}{}}{
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
     \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-  \draw #2 rectangle #3 node[pos=.5] {\footnotesize #1};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1> ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw<#1>[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+
+  }
 }
 
-\newcommand{\actsf}[3][]{
-  \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+\newcommand{\actsf}[#4][]{
+  \ifthenelse{\equal{#1}{}}{
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {#4};
+  }{
+   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
+ }
 }
@@
 \newcommand{\inactsf}[3][]{
+  \ifthenelse{\equal{#1}{}}{
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-  \draw #2 rectangle #3 node[pos=.5] {\footnotesize #1};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }
 }
@@
     % attacker calls trusted code again
-    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{onlyenv}<+->
+      \inactadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
     \begin{onlyenv}<.>
@@
     % attacker calls trusted code again
-    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{onlyenv}<+->
+    \inactadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
     \begin{onlyenv}<.-.(1)>

### 2017-11-30 presentations/illustrations/presentation-illustrations.tex
Added small example with linear capability, no attack

@@
 
+% Active adversary
 \newcommand{\actadv}[4][]{
@@
   }{
-    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw<#1>[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-
+    \onlyenv<#1>
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \endonlyenv
   }
@@
 
+% Inactive adversary
 \newcommand{\inactadv}[4][]{
@@
   }{
-    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw<#1>[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-
+    \onlyenv<#1>
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
+    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \endonlyenv
   }
@@
 
-\newcommand{\actsf}[#4][]{
+% Active stack frame
+\newcommand{\actsf}[4][]{
   \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {#4};
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
   }{
@@
 
-\newcommand{\inactsf}[3][]{
+% Inactive stack frame
+\newcommand{\inactsf}[4][]{
   \ifthenelse{\equal{#1}{}}{
@@
   }
+}
+
+\newcommand{\capbrace}[3][sp1]{
+  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
+
 }
@@
 
+\begin{frame}
+    \frametitle{Linear caps: stack pointers}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw (0,0) -- (0,15);
+      \draw (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % Caller 
+    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+    \begin{onlyenv}<+>
+      \capbrace{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % Split the stack pointer
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    % Callee called
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[fill=gray!50, right=0cm of sp2] { private stack pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actsf{(0,5)}{(6,6.5)}{callee stack frame}
+    \end{onlyenv}
+
+    % Return from callee
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    % Caller splices the stack frames
+    \begin{onlyenv}<+>
+      \capbrace{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+  \end{tikzpicture}
+\end{frame}
+
 \end{document}

### 2017-12-01 presentations/illustrations/presentation-illustrations.tex
Added a series og slides illustrating how linear capabilities prevent attack 1

@@
 \usepackage{tikzpeople}
+\usepackage{amssymb}
 
@@
     % attacker...
-    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller satck frame};
+    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
     \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
@@
     \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
-
     \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
@@
       \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { private stack pointer };
+      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
       \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
@@
 
+\begin{frame}
+    \frametitle{Linear caps prevent attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \draw (0,0) -- (0,15);
+      \draw (6,0) -- (6,15);
+      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \end{scope}
+
+    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+    % Caller 
+    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+    \begin{onlyenv}<+>
+      \capbrace{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % Split the stack pointer
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    % Callee called
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+
+    \begin{onlyenv}<.->
+      \begin{scope}
+        \clip (13.9,9) rectangle (18.1,14);
+        \draw[fill=gray!20] (14,8.9) rectangle (18,14.1);
+      \end{scope}
+    \end{onlyenv}
+    \draw<.>[fill=white] (14,10) rectangle node[color=red] {} (18,10.5);
+
+    % Callee stores stack pointer
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (8,5) -- (8,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
+      \draw (16,15) node {heap memory};
+    \end{onlyenv}
+    \begin{onlyenv}<.>
+      \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
+      \draw node[right=0cm of usp] { stack pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+    \end{onlyenv}
+
+    % Return from callee
+    \begin{onlyenv}<+->
+      \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
+      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[right=0cm of sp2] { private stack pointer };
+      \draw node[right=0cm of usp] { stack pointer };
+    \end{onlyenv}
+
+    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
+    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
+    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};
+
+  \end{tikzpicture}  
+\end{frame}
+
 \end{document}

### 2017-12-01 presentations/illustrations/presentation-illustrations.tex
Added slides for linear capabilities and attack 2. Cleaned up some tikz code.

@@
 
+}
+
+\newcommand{\stdstackstart}{
+  \scope
+    \clip (-.1,-.1) rectangle (6.1,15.1);
+    \fill[fill=white] (0,0) rectangle (6,15);
+    \draw (0,0) -- (0,15);
+    \draw (6,0) -- (6,15);
+    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+  \endscope
+  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
 }
@@
       \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
       \draw[draw=gray!80] (0,0) -- (0,15);
@@
     % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (6.1,15);
-        \draw[fill=white] (0,2.5) rectangle (6,15.1);
-      \end{scope}
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \stdstackstart
 
@@
       \begin{scope}
-        \clip (13.9,6) rectangle (18.1,13);
-        \draw[fill=gray!20] (14,5.9) rectangle (18,13.1);
+        \fill[gray!20] (14,6) rectangle (18,13);
+        \draw (14,6) -- (14,13);
+        \draw (18,6) -- (18,13);
       \end{scope}
@@
     % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \draw[fill=white] (0,4) rectangle (6,16);
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \stdstackstart
 
@@
     % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \draw[fill=white] (0,9.5) rectangle (6,15.2);
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \stdstackstart
 
@@
     \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw[fill=gray!20] (14,5.9) rectangle (18,13.1);
+      \fill[gray!20] (14,6) rectangle (18,13);
+      \draw (14,6) -- (14,13);
+      \draw (18,6) -- (18,13);
     \end{scope}
@@
     % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw (0,0) -- (0,15);
-      \draw (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \stdstackstart
 
@@
       \begin{scope}
-        \clip (13.9,9) rectangle (18.1,14);
-        \draw[fill=gray!20] (14,8.9) rectangle (18,14.1);
+        \fill[gray!20] (14,9) rectangle (18,14);
+        \draw (14,9) -- (14,14);
+        \draw (18,9) -- (18,14);
       \end{scope}
@@
     \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
+  \end{tikzpicture}  
+\end{frame}
+
+\begin{frame}
+  \frametitle{Linear caps prevents attack 2}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % Caller 
+    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+    \begin{onlyenv}<+>
+      \capbrace{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % Split the stack pointer
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    % Callee called
+    \begin{onlyenv}<+>
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+
+    \begin{onlyenv}<+-.(1)>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+
+      \capbrace[usp]{(13.5,5)}{(13.5,16)}
+      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+
+      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+
+    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
+
+    \begin{onlyenv}<+->
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[right=0cm of sp2] { return data pointer };
+
+      \capbrace[usp]{(13.5,5)}{(13.5,16)}
+      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+
+      \begin{scope}
+        \fill[fill=gray!50] (0,2.5) rectangle (6,15) node[pos=.5,color=black] {};
+        \draw (0,0) -- (0,15);
+        \draw (6,0) -- (6,15);
+      \end{scope}
+      \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
+      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+  \end{onlyenv}
+
+  \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
+  \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
+  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { Splice fails, no\\stack pointer returned.};
 

### 2017-12-01 presentations/illustrations/presentation-illustrations.tex
A couple of updates to the new bit

@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
       \draw node[right=0cm of usp] { stack pointer };
@@
     \begin{onlyenv}<+>
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
@@
 
-    \begin{onlyenv}<+-.(1)>
+    \begin{onlyenv}<+-.(2)>
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
@@
 
-    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
+    % At this point we have a linear point stored in a part of memory that it governs.
+In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
+    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
 

### 2017-12-04 presentations/illustrations/presentation-illustrations.tex
adjusted the linear capability prevents attack 2 slides to fit directly with attack 2.

@@
     \stdstackstart
+    \draw (16.5,5.5) node {register file};
+    \begin{scope}
+      \clip (15.9,1.1) rectangle (17.1,5.1);
+      \foreach \x in {1,2,...,4}
+      {
+        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
+      };
+    \end{scope}
+    \node (r1) at (16.75,4.5) {};
+    \node (r2) at (16.75,3.5) {};
 
@@
     \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
+      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
       \draw node[right=0cm of sp1] { stack pointer };
+      \draw (r1) edge[->,red,bend left] (sp1);
     \end{onlyenv}
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw (r2) edge[->,red,bend left] (sp2);
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw (r1) edge[->,red,bend left] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw (r2) edge[->,red,bend left] (sp2);
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw (r1) edge[->,red,bend left] (usp);
       \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { return data pointer };
       \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
@@
 
-    \begin{onlyenv}<+-.(2)>
+    \begin{onlyenv}<+-.(1)>
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
@@
 In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
-    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
+%    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
 
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of sp2] { return data pointer };
+      \draw node[fill=gray!50,right=0cm of sp2] { return data pointer };
 
@@
       \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \inactadv{(0,5)}{(6,8)}{callee stack frame}
+      \actsf{(0,8)}{(6,10.5)}{callee (2) stack frame} 
   \end{onlyenv}
 
-  \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
-  \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
-  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { Splice fails, no\\stack pointer returned.};
+  \draw<+>[teal,very thick] (-0.5,7.5) -- (6.5,11);
+  \draw<.>[teal,very thick] (-0.5,11) -- (6.5,7.5);
+  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { No stack pointer\\for callee (2)!};
 

### 2017-12-05 presentations/illustrations/presentation-illustrations.tex
finished register file part for attack 1.

@@
       \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,bend left] (sp1);
+      \draw (r1) edge[->,red,thick,bend left] (sp1);
     \end{onlyenv}
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
+      \draw (r2) edge[->,red,thick,bend left] (sp2);
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
+      \draw (r1) edge[->,red,thick,bend left] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
       \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { return data pointer };
+
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
+      \draw node[right=0cm of sp2] { return data pointer };
       \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
@@
     \end{onlyenv}
+    \draw<.-> (r2) edge[->,red,bend left] (sp2);
 
@@
     \end{onlyenv}
+
     \begin{onlyenv}<.>
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
+      \draw node[right=0cm of sp2] { return data pointer };
     \end{onlyenv}

### 2017-12-05 fully-abstract-comp/technical-report.tex
removed memory from halted configuration.
Comments and small fixes.

@@
     \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
-    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
+    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus \{\halted\}
   \end{array}
@@
 $i$ defines the set $\Instr$
+
+\dominique{4-12-2017: some instructions need to be able to take an integer rather than a natural, e.g. cca (as used in the implementation of call).}
 
@@
 In fact, $\scall{o}{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
-We therefore introduce $\scall[1]{o}{r_1}{r_2},\dots,\scall[\calllen-1]{o}{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{o}{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
+We therefore introduce $\scall[0]{o}{r_1}{r_2},\dots,\scall[\calllen-1]{o}{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{o}{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
 We define the following condition that indicates that $\aaddr$ is the first address of a $\scall{o}{r_1}{r_2}$ instruction in the configuration $\Phi$:
@@
       \begin{array}{l}
-        \Phi.\mem(\aaddr) = \scall[1]{o}{r_1}{r_2} \tand\\
+        \Phi.\mem(\aaddr) = \scall[0]{o}{r_1}{r_2} \tand\\
         \vdots \\
@@
     \tstore{\rstk}{42}\\
-    \tcca{\rstk}{-1}\\
+    \tcca{\rstk}{(-1)}\\
     \text{// split the stack at its current address - rstk done.}\\
@@
 }
+\dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
 The call code does the following:
@@
 
+\dominique{4-12-2017: what about $\pi_a$?}
+
 \[
@@
   \sem{\tfail}(\Phi) = & \; \failed \\
-  \sem{\thalt}(\Phi) = & \; (\halted, \Phi.\mem)
+  \sem{\thalt}(\Phi) = & \; \halted
 \end{align*}
@@
 \lau{15-09-2017: Should gete and getb do something for seals?}
+\dominique{4-12-2017: doesn't matter, I think.}
 
@@
 \lau{15-09-2017: Should getl do something for seals?}
+\dominique{4-12-2017: doesn't matter, I think.}
 \lau{15-09-2017: Do we want to allow getl and getp to work on sealed capabilities?}
+\dominique{Linearity of sealed caps is already observable anyway, so no point in hiding it.
+Permissions of the sealed capability should be hidden, I think, mostly out of conservativity (hide as much as possible!).}
+\dominique{4-12-2017: Couldn't the definitions above use $\pi_b$, $\pi_e$, $\pi_l$ and $\pi_a$ and $\pi_p$ (the latter two to be added) ?}
 
@@
 \noindent\textbf{cca}\\
+\dominique{4-12-2017: The phrase $\text{either $n = \rn$ or $n = \Phi.\reg(\rn)$ and in either case $n \in \ints$ and}$ is repeated a lot in the semantics.
+Perhaps define an abbreviation as $\Phi(\rn) = n$?}
+
 \emph{Change Current Address}
@@
 \end{align*}
+\dominique{4-12-2017: $e_\stk$ is no longer stored in $\ms_{\stk,\priv}$ in updated call implementation.}
+
+Note: the caller may have split part of the stack pointer off and even pass the fragments split off to the callee in registers.
+This behavior is in principle fine.
+Source semantics will define that only the non-split-off part will be encapsulated.
+The parts that were split off and passed to the adversary are not protected, as expected.
 
@@
   \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
-  \item $\seal{\_,\vsc}$ and $\vsc$ respects write-xor-execute
+  \item $\sealed{\_,\vsc}$ and $\vsc$ respects write-xor-execute
   \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
@@
   \lau{27-09-2017: The ``somehow''s need to be fixed.}
-  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{\sigma,c_\code}$ and $\vsc_\data = \sealed{\sigma,c_\data}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
+  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{\sigma,c_\code}$ and $\vsc_\data = \sealed{\sigma,c_\data}$, and two memory segments $\ms$ and $\ms_\link$ satisfying
   \begin{itemize}
@@
 \]
-$\oplus$ is defined as follows
+$\oplus$ on regions is defined as follows
 \begin{align*}
@@
 \lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
+\dominique{4-12-2017: Regarding the question whether $\opc$ should be mentioned in the memory satisfaction relation.
+  The relevant intuitive aspect of the proof here, I think, is that on a return, the source semantics restores opc from the copy kept in the stack, while the target semantics restores it from the copy in the code part of the return pointer.
+  We will need to prove (somewhere) that these two correspond, which means we will need to track of the invariant that guarantees this somewhere (perhaps here?).
+  I suspect the relevant invariants are (1) that any executable pointer signed with $\sealAss{a}$ must point to $a + \calllen$ and (2) that anything related to a $\rretc$ must be executable and sealed.
+  Perhaps this invariant is easiest to track in (1) the general requirement on the $H_\sigma$ and (2) in the value relation.
+  This suggests nothing should be required about opc here?
+  To be confirmed in the proof of the fundamental theorem.
+}
 
@@
     \begin{array}{l}
-      \exists S : \RegName \fun \World \ldotp \\
-      \quad W = \bigoplus_{r \in \RegName} S(r) \\
+      \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
+      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg \} \cup R))} S(r) \\
       \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\

### 2017-12-07 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex

@@
-\documentclass[sigplan, review]{acmart}
+\documentclass[sigplan,screen]{acmart}
 
@@
 % Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
 \setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-% \acmDOI{10.475/123_4}
-
-% ISBN
-% \acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[PriSC]{Workshop on Principles of Secure Compilation}{January 2018}{Los Angeles, California USA} 
+\acmDOI{}
+\acmISBN{}
+\acmConference[PriSC]{Workshop on Principles of Secure Compilation}{2018}{Los Angeles} 
 \acmYear{2018}
 \copyrightyear{2018}
+\settopmatter{printacmref=false}
 
@@
 We propose and study a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
-The calling convention is based on linear capabilities, a new type of capabilities that are prevented by the hardware from being duplicated.
+The calling convention is based on linear capabilities, a type of capabilities that are prevented by the hardware from being duplicated.
 In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, building on the concept of fully abstract compilation.
@@
 An alternative that does not have this requirement, relies on \emph{linear capabilities}, rather than local ones.
-Linear capabilities are (to the best of our knowledge) a new concept, similar to CHERI's local capabilities.\footnote{One author of this abstract is also investigating linear capabilities in another application, also submitted to PriSC.}
+Although details in the literature are scarce, linear capabilities have already been implemented in the SAFE machine~\citep{amorim_verified_2014}\footnote{Private communication with Catalin Hrit\c{c}u.} and considered for implementation in CHERI\footnote{Private communication with Robert Watson.}.
 The idea is that all capabilities are marked as either regular or linear.

### 2017-12-12 fully-abstract-comp/technical-report.tex
Answered Dominique's comments and made a few related updates

@@
 \begin{array}{rcl}
-n &\in & \nats \\
+n &\in & \ints \\
 \trg{r} &\in &  \tRegName \\
@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{r}}{\trg{r}} \mid \\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
   & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
@@
 
-\subsubsection{Helpful functions and sets}
+\subsubsection{Helpful functions, sets, and conventions}
 
@@
 \]
+
 \[
@@
 \dominique{4-12-2017: what about $\pi_a$?}
-
+\lau{12-12-2017: I have added them as I needed them in the LR. I have not needed $\pi_a$ yet.}
 \[
@@
 
+For $\rn$ which can be a register or an integer, we take
+\[
+  \Phi(\rn) = n
+\]
+to mean
+\[
+  \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$ and in either case $n \in \ints$}
+\]
+
 \subsubsection{Step Relation}
@@
   \sem{\tfail}(\Phi) = & \; \failed \\
-  \sem{\thalt}(\Phi) = & \; \halted
+  \sem{\thalt}(\Phi) = & \; \haltde
 \end{align*}
@@
 \dominique{4-12-2017: Couldn't the definitions above use $\pi_b$, $\pi_e$, $\pi_l$ and $\pi_a$ and $\pi_p$ (the latter two to be added) ?}
+\lau{12-12-2017: $\pi_\_$ also projects from sealed capabilities, so the case of sealed capabilities would have to be explicitely excluded (with getl as the exception).}
 
@@
                                     \begin{array}[t]{l}
-                                      \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
-                                      \quad\text{ and in either case $n \in \ints$ and} \\
+                                      \Phi(\rn) = n \tand \\
                                       \Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
@@
                                     \begin{array}[t]{l}
-                                      \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
-                                      \quad\text{ and in either case $n \in \ints$ and} \\
+                                      \Phi(\rn) = n \tand \\
                                       \Phi(r) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
@@
                                     \begin{array}[t]{l}
-                                      \text{either $n = \rn$ or $n = \Phi(\rn)$}\\
-                                      \quad\text{ and in either case $n \in \ints$ and} \\
+                                      \Phi(\rn) = n\tand \\
                                       \Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
@@
                                           \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                          \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
-                                          \quad\text{and in either case $n \in \ints$ and} \\
+                                          \Phi(\rn) = n \tand\\
                                           \decPerm{n} \sqsubseteq \perm \tand \\
@@
                                           \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                          \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$}\\
-                                          \quad\text{and in either case $n \in \ints$ and} \\
+                                          \Phi(\rn) = n \tand\\
                                           \decPerm{n} \sqsubseteq \perm \tand \\
@@
 \begin{align*}
-  \sem{\tlt{r_0}{r_1}{r_2}}(\Phi) = &
+  \sem{\tlt{r_0}{\rn_1}{\rn_2}}(\Phi) = &
                                                   \begin{cases}
@@
                                                       \text{if for $i \in \{1,2\}$}\\
-                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
-                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                      \quad\Phi(\rn_i) = n_i \tand\\
                                                       \quad\text{and $n_1 < n_2$}\\        
@@
                                                       \text{if for $i \in \{1,2\}$}\\
-                                                      \quad\text{$n_i = r_i$ or $n_i = \Phi(r_i)$}\\
-                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                      \quad \Phi(\rn_i) = n_i \tand \\
                                                       \quad\text{and $n_1 \not< n_2$}\\        
@@
                                                       \text{if for $i \in \{1,2\}$}\\
-                                                      \quad\text{$n_i = \rn_i$ or $n_i = \Phi(\rn_i)$}\\
-                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                      \quad \Phi(\rn_i) = n_i\\
                                                     \end{array}\\
@@
                                                       \text{if for $i \in \{1,2\}$}\\
-                                                      \quad\text{$n_i = \rn_i$ or $n_i = \Phi(\rn_i)$}\\
-                                                      \quad\text{and in either case $n_i \in \ints$}\\
+                                                      \quad\Phi(\rn_i) = n_i\\
                                                     \end{array}\\
@@
                                    \Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                   \Phi(\rn_4) = n \tor \rn_4 = n \\
-                                   \quad\text{ and in either case } n \in \nats \\
+                                   \Phi(\rn_4) = n \tand\\
                                    \baddr \leq n \tand n < \eaddr \tand \\
@@
                                    \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
-                                   \Phi(\rn_4) = n \tor \rn_4 = n \\
-                                   \quad\text{ and in either case } n \in \nats \\
+                                   \Phi(\rn_4) = n \tand\\
                                    \sigma_\baddr \leq n \tand n < \sigma_\eaddr \tand \\
@@
                                    \Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
-                                   \Phi(\rn_4) = n \tor \rn_4 = n \\
-                                   \quad\text{ and in either case } n \in \nats \\
+                                   \Phi(\rn_4) = n \tand \\
                                    \baddr \leq n \tand n < \eaddr \tand \\
@@
                                                    \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
-                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{\eaddr_\stk} \tand\\
+                                                   \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} \tand\\
                                                    \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk - 1]} \tand \\
@@
 \dominique{4-12-2017: $e_\stk$ is no longer stored in $\ms_{\stk,\priv}$ in updated call implementation.}
+\lau{12-12-2017: Correct. Please feel free to correct these things.}
 

### 2017-12-13 presentations/illustrations/presentation-illustrations.tex
Added register file to attack 2, linear capabilities.
Added an attack 3 and the linear case

@@
 \begin{frame}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \frametitle{Stack and return capabilities: Attack 3}
+    \begin{scope}
+    \clip (-.1,-.1) rectangle (5.1,10.1);
+    \fill[fill=gray!50] (0,0) rectangle (5,10);
+    \draw (0,0) -- (0,10);
+    \draw (5,0) -- (5,10);
+    \end{scope}
+    \draw (2.5,11) node {heap memory};
+
+    \begin{scope}
+    \clip (12.9,-.1) rectangle (18.1,11.1);
+    \fill[fill=gray!50] (13,0) rectangle (18,11);
+    \draw (13,0) -- (13,11);
+    \draw (18,0) -- (18,11);
+    \end{scope}
+    \draw (15.5,12) node {more heap memory};
+
+    \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+
+    \begin{onlyenv}<+->
+      \capbrace[ret1]{(5.1,1)}{(5.1,4)}
+      \draw node[fill=gray!50, right=0cm of ret1] { \phantom{ret-ptr} };
+      \draw node[right=0cm of ret1] { ret-ptr };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
+    \end{onlyenv}
+
+    \begin{onlyenv}<.-.(2)>
+      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
+      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
+      \draw (18,4.75)  edge[->,bend left] (copy);
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+      \draw (13,1.25) edge[->,red,thick,bend left] (ret1);
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      (12.9,3) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
+      \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
+      \draw node[left=0cm of ret2] { ret-ptr (1) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=gray!30] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \capbrace[ret3]{(5.1,5)}{(5.1,8)}
+      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+      \draw node[right=0cm of ret3] { ret-ptr (2) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw (21,7.5) node[devil,opaque=0.5,mirrored,minimum size=1cm] {};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
+      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
+      \draw (18,8.25)  edge[->,red,thick,bend left] (copy);
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
+      \draw[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw (9,5.5) node [ellipse, draw, red, thick,minimum height=1.5cm, minimum width=3cm] { };
+      \draw node[red,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  
+    \end{onlyenv}
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
   \frametitle{Local Stack Capabilities prevent Attack 1}
@@
 \begin{frame}
+  \frametitle{Local Stack Capabilities prevent Attack 3}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    \begin{scope}
+      \fill[gray!50] (13,0) rectangle (18,13);
+      \draw (13,0) -- (13,13);
+      \draw (18,0) -- (18,13);
+      \draw (15.5,14) node {heap memory};
+    \end{scope}
+
+    \begin{onlyenv}<+->
+      \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \end{onlyenv}
+
+% Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,2.5) rectangle (8.1,5.5);
+        \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
+      \draw node[right=0cm of ret1] { ret-ptr };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \end{onlyenv}
+
+    \begin{onlyenv}<+-.(2)>
+      \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
+    \end{onlyenv}
+
+    \begin{onlyenv}<+>
+      \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
+      \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
+      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \end{onlyenv}
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,5.5) rectangle (8.1,8.5);
+        \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
+      \draw node[right=0cm of ret2] { ret-ptr (1) };
+    \end{onlyenv}
+    \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \foreach \x in {11.5,12,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,8.5) rectangle (8.1,11.5);
+        \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+      \draw node[right=0cm of ret3] { ret-ptr (2) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+      \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
+    \end{onlyenv}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
     \frametitle{Linear caps: stack pointers}
@@
     \node (r1) at (16.75,4.5) {};
-    \node (r2) at (16.75,3.5) {};
+    \node (r2) at (16.75,3.6) {};
 
@@
       (8,5) -- (8,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
       \draw (16,15) node {heap memory};
@@
     \begin{onlyenv}<.>
+      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
       \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
@@
     \begin{onlyenv}<+->
+      \draw (14,10.25) edge[->,red,bend right] (usp);
       \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
@@
 
+    \draw (16.5,5.5) node {register file};
+    \begin{scope}
+      \clip (15.9,1.1) rectangle (17.1,5.1);
+      \foreach \x in {1,2,...,4}
+      {
+        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
+      };
+    \end{scope}
+    \node (r1) at (16.75,4.5) {};
+    \node (r2) at (16.75,3.6) {};
+
     % Caller 
@@
       \draw node[right=0cm of sp1] { stack pointer };
+      \draw (r1) edge[->,red,thick,bend left] (sp1);
     \end{onlyenv}
@@
       \capbrace[usp]{(6.5,5)}{(6.5,16)}
+      \draw (r2) edge[->,red,thick,bend left] (sp2);
+      \draw (r1) edge[->,red,thick,bend left] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw (r1) edge[->,red,bend left] (usp);
       \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
       \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
@@
 
-    \begin{onlyenv}<+-.(1)>
+    \begin{onlyenv}<.->
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
+      \draw node[fill=gray!50,right=0cm of sp2] { \phantom{return data pointer} };
+      \draw (r2) edge[->,red,bend left] (sp2);
+      \draw node[right=0cm of sp2] { return data pointer };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+>
 
@@
     \begin{onlyenv}<+->
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50,right=0cm of sp2] { return data pointer };
 

### 2017-12-13 presentations/illustrations/attacks.tex
Restructured document to allow for better compile time during editing.

--- 2017-12-04 presentations/illustrations/presentation-illustrations.tex
@@
-\documentclass{beamer}
-
-\usepackage{ifthen}
-\usepackage{pgf}
-\usepackage{ulem}
-\usepackage{tikz}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{calc}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
-\usepackage{amssymb}
-
-% Active adversary
-\newcommand{\actadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
-
-% Inactive adversary
-\newcommand{\inactadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
-
-% Active stack frame
-\newcommand{\actsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
- }
-}
-
-% Inactive stack frame
-\newcommand{\inactsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }
-}
-
-\newcommand{\capbrace}[3][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
-
-}
-
-\newcommand{\stdstackstart}{
-  \scope
-    \clip (-.1,-.1) rectangle (6.1,15.1);
-    \fill[fill=white] (0,0) rectangle (6,15);
-    \draw (0,0) -- (0,15);
-    \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-  \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-}
-
-% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
-\makeatletter
-\newif\if@showgrid@grid
-\newif\if@showgrid@left
-\newif\if@showgrid@right
-\newif\if@showgrid@below
-\newif\if@showgrid@above
-\tikzset{%
-    every show grid/.style={},
-    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
-    show grid/.default={true},
-    show grid/.cd,
-    labels/.style={font={\sffamily\small},help lines},
-    xlabels/.style={},
-    ylabels/.style={},
-    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
-    true/.style={left,below},
-    false/.style={left=false,right=false,above=false,below=false,grid=false},
-    none/.style={left=false,right=false,above=false,below=false},
-    all/.style={left=true,right=true,above=true,below=true},
-    grid/.is if=@showgrid@grid,
-    left/.is if=@showgrid@left,
-    right/.is if=@showgrid@right,
-    below/.is if=@showgrid@below,
-    above/.is if=@showgrid@above,
-    false,
-}
-
-\def\@showgrid#1{%
-    \begin{scope}[every show grid,show grid/.cd,#1]
-    \if@showgrid@grid
-    \begin{pgfonlayer}{background}
-    \draw [help lines]
-        (current bounding box.south west) grid
-        (current bounding box.north east);
-%
-    \pgfpointxy{1}{1}%
-    \edef\xs{\the\pgf@x}%
-    \edef\ys{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{south west}
-    \edef\xa{\the\pgf@x}%
-    \edef\ya{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{north east}
-    \edef\xb{\the\pgf@x}%
-    \edef\yb{\the\pgf@y}%
-    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
-    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
-    \if@showgrid@below
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
-    }
-    \fi
-    \if@showgrid@above
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
-    }
-    \fi
-    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
-    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
-    \if@showgrid@left
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
-    }
-    \fi
-    \if@showgrid@right
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
-    }
-    \fi
-    \end{pgfonlayer}
-    \fi
-    \end{scope}
-}
-\makeatother
-
-\begin{document}
-\begin{frame}
-  \frametitle{Traditional stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
-    \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
-
-    % traditional:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-
-    % traditional return:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call, attack:
-    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
-
-    % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
-
-    % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
-      skip caller frame};
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+-+(4)>
-      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(3)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    %prepare call: store old return pointer
-    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1); 
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
-      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { new stack pointer };
-    \end{onlyenv}
-
-    % call using capabilities
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
 \begin{frame}
@@
 \begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 1}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
+    \frametitle{Stack and return capabilities: Attack 3}
+    \begin{scope}
+    \clip (-.1,-.1) rectangle (5.1,10.1);
+    \fill[fill=gray!50] (0,0) rectangle (5,10);
+    \draw (0,0) -- (0,10);
+    \draw (5,0) -- (5,10);
+    \end{scope}
+    \draw (2.5,11) node {heap memory};
 
-    % attacker...
-    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
-    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw node[right=0cm of sp1] { stack pointer };
+    \begin{scope}
+    \clip (12.9,-.1) rectangle (18.1,11.1);
+    \fill[fill=gray!50] (13,0) rectangle (18,11);
+    \draw (13,0) -- (13,11);
+    \draw (18,0) -- (18,11);
+    \end{scope}
+    \draw (15.5,12) node {more heap memory};
 
-    % stores stack pointer in heap
-    \begin{scope}
-      \fill[gray!20] (14,6) rectangle (18,13);
-      \draw (14,6) -- (14,13);
-      \draw (18,6) -- (18,13);
-    \end{scope}
-    \draw (16,14) node {heap memory};
-    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
-    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
+    \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+
+    \begin{onlyenv}<+->
+      \capbrace[ret1]{(5.1,1)}{(5.1,4)}
+      \draw node[fill=gray!50, right=0cm of ret1] { \phantom{ret-ptr} };
+      \draw node[right=0cm of ret1] { ret-ptr };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
+    \end{onlyenv}
+
+    \begin{onlyenv}<.-.(2)>
+      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
+      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
+      \draw (18,4.75)  edge[->,bend left] (copy);
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+      \draw (13,1.25) edge[->,red,thick,bend left] (ret1);
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      (12.9,3) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
+      \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
+      \draw node[left=0cm of ret2] { ret-ptr (1) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=gray!30] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \capbrace[ret3]{(5.1,5)}{(5.1,8)}
+      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+      \draw node[right=0cm of ret3] { ret-ptr (2) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw (21,7.5) node[devil,opaque=0.5,mirrored,minimum size=1cm] {};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
+      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
+      \draw (18,8.25)  edge[->,red,thick,bend left] (copy);
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
+      \draw[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw (9,5.5) node [ellipse, draw, red, thick,minimum height=1.5cm, minimum width=3cm] { };
+      \draw node[red,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  
+    \end{onlyenv}
   \end{tikzpicture}
 \end{frame}
-
-\begin{frame}
-  \frametitle{Local caps: Stack clearing prevents Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);
-
-    % attacker calls trusted code again
-    \begin{onlyenv}<+->
-      \inactadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % trusted code clears the stack
-    \begin{onlyenv}<+->
-      \foreach \x in {9,9.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
-    (13,13.5) -- (10,14.5);
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.>
-      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker returns
-    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % trusted code clears stack again
-    \begin{onlyenv}<+->
-      \foreach \x in {7,7.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    % then, the trusted code can return
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Linear caps: stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actsf{(0,5)}{(6,6.5)}{callee stack frame}
-    \end{onlyenv}
-
-    % Return from callee
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Caller splices the stack frames
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Linear caps prevent attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-    \draw (16.5,5.5) node {register file};
-    \begin{scope}
-      \clip (15.9,1.1) rectangle (17.1,5.1);
-      \foreach \x in {1,2,...,4}
-      {
-        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
-      };
-    \end{scope}
-    \node (r1) at (16.75,4.5) {};
-    \node (r2) at (16.75,3.5) {};
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,bend left] (sp1);
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<.->
-      \begin{scope}
-        \fill[gray!20] (14,9) rectangle (18,14);
-        \draw (14,9) -- (14,14);
-        \draw (18,9) -- (18,14);
-      \end{scope}
-    \end{onlyenv}
-    \draw<.>[fill=white] (14,10) rectangle node[color=red] {} (18,10.5);
-
-    % Callee stores stack pointer
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (8,5) -- (8,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
-      \draw (16,15) node {heap memory};
-    \end{onlyenv}
-    \begin{onlyenv}<.>
-      \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
-      \draw node[right=0cm of usp] { stack pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-    \end{onlyenv}
-
-    % Return from callee
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
-      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of sp2] { private stack pointer };
-      \draw node[right=0cm of usp] { stack pointer };
-    \end{onlyenv}
-
-    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
-    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
-    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
-  \end{tikzpicture}  
-\end{frame}
-
-\begin{frame}
-  \frametitle{Linear caps prevents attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<+-.(1)>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    % At this point we have a linear point stored in a part of memory that it governs.
-In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
-%    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
-
-    \begin{onlyenv}<+->
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50,right=0cm of sp2] { return data pointer };
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \begin{scope}
-        \fill[fill=gray!50] (0,2.5) rectangle (6,15) node[pos=.5,color=black] {};
-        \draw (0,0) -- (0,15);
-        \draw (6,0) -- (6,15);
-      \end{scope}
-      \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \inactadv{(0,5)}{(6,8)}{callee stack frame}
-      \actsf{(0,8)}{(6,10.5)}{callee (2) stack frame} 
-  \end{onlyenv}
-
-  \draw<+>[teal,very thick] (-0.5,7.5) -- (6.5,11);
-  \draw<.>[teal,very thick] (-0.5,11) -- (6.5,7.5);
-  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { No stack pointer\\for callee (2)!};
-
-  \end{tikzpicture}
-\end{frame}
-
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "presentation-illustrations"
+%%% End:

### 2017-12-13 presentations/illustrations/linear-prevention.tex
Restructured document to allow for better compile time during editing.

--- 2017-12-13 presentations/illustrations/presentation-illustrations.tex
@@
-\documentclass{beamer}
-
-\usepackage{ifthen}
-\usepackage{pgf}
-\usepackage{ulem}
-\usepackage{tikz}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{calc}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
-\usepackage{amssymb}
-
-% Active adversary
-\newcommand{\actadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
-
-% Inactive adversary
-\newcommand{\inactadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
-
-% Active stack frame
-\newcommand{\actsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
- }
-}
-
-% Inactive stack frame
-\newcommand{\inactsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }
-}
-
-\newcommand{\capbrace}[3][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
-
-}
-
-\newcommand{\stdstackstart}{
-  \scope
-    \clip (-.1,-.1) rectangle (6.1,15.1);
-    \fill[fill=white] (0,0) rectangle (6,15);
-    \draw (0,0) -- (0,15);
-    \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-  \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-}
-
-% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
-\makeatletter
-\newif\if@showgrid@grid
-\newif\if@showgrid@left
-\newif\if@showgrid@right
-\newif\if@showgrid@below
-\newif\if@showgrid@above
-\tikzset{%
-    every show grid/.style={},
-    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
-    show grid/.default={true},
-    show grid/.cd,
-    labels/.style={font={\sffamily\small},help lines},
-    xlabels/.style={},
-    ylabels/.style={},
-    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
-    true/.style={left,below},
-    false/.style={left=false,right=false,above=false,below=false,grid=false},
-    none/.style={left=false,right=false,above=false,below=false},
-    all/.style={left=true,right=true,above=true,below=true},
-    grid/.is if=@showgrid@grid,
-    left/.is if=@showgrid@left,
-    right/.is if=@showgrid@right,
-    below/.is if=@showgrid@below,
-    above/.is if=@showgrid@above,
-    false,
-}
-
-\def\@showgrid#1{%
-    \begin{scope}[every show grid,show grid/.cd,#1]
-    \if@showgrid@grid
-    \begin{pgfonlayer}{background}
-    \draw [help lines]
-        (current bounding box.south west) grid
-        (current bounding box.north east);
-%
-    \pgfpointxy{1}{1}%
-    \edef\xs{\the\pgf@x}%
-    \edef\ys{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{south west}
-    \edef\xa{\the\pgf@x}%
-    \edef\ya{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{north east}
-    \edef\xb{\the\pgf@x}%
-    \edef\yb{\the\pgf@y}%
-    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
-    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
-    \if@showgrid@below
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
-    }
-    \fi
-    \if@showgrid@above
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
-    }
-    \fi
-    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
-    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
-    \if@showgrid@left
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
-    }
-    \fi
-    \if@showgrid@right
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
-    }
-    \fi
-    \end{pgfonlayer}
-    \fi
-    \end{scope}
-}
-\makeatother
-
-\begin{document}
-\begin{frame}
-  \frametitle{Traditional stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
-    \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
-
-    % traditional:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-
-    % traditional return:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call, attack:
-    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
-
-    % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
-
-    % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
-      skip caller frame};
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+-+(4)>
-      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(3)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    %prepare call: store old return pointer
-    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1); 
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
-      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { new stack pointer };
-    \end{onlyenv}
-
-    % call using capabilities
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker...
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % stores stack pointer in heap
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \fill[gray!20] (14,6) rectangle (18,13);
-        \draw (14,6) -- (14,13);
-        \draw (18,6) -- (18,13);
-      \end{scope}
-      \draw (16,14) node {heap memory};
-      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \end{onlyenv}
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
-    \end{onlyenv}
-
-    % attacker returns
-    \begin{onlyenv}<+>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-
-    % attacker gets called again + uses old stack pointer.
-    \begin{onlyenv}<+>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
-      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+>
-      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<.-.(2)>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
-    \end{onlyenv}
-
-    % attacker calls trusted code again
-%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{onlyenv}<+->
-      \inactadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.-.(1)>
-      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    \frametitle{Stack and return capabilities: Attack 3}
-    \begin{scope}
-    \clip (-.1,-.1) rectangle (5.1,10.1);
-    \fill[fill=gray!50] (0,0) rectangle (5,10);
-    \draw (0,0) -- (0,10);
-    \draw (5,0) -- (5,10);
-    \end{scope}
-    \draw (2.5,11) node {heap memory};
-
-    \begin{scope}
-    \clip (12.9,-.1) rectangle (18.1,11.1);
-    \fill[fill=gray!50] (13,0) rectangle (18,11);
-    \draw (13,0) -- (13,11);
-    \draw (18,0) -- (18,11);
-    \end{scope}
-    \draw (15.5,12) node {more heap memory};
-
-    \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-
-    \begin{onlyenv}<+->
-      \capbrace[ret1]{(5.1,1)}{(5.1,4)}
-      \draw node[fill=gray!50, right=0cm of ret1] { \phantom{ret-ptr} };
-      \draw node[right=0cm of ret1] { ret-ptr };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=white] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
-    \end{onlyenv}
-
-    \begin{onlyenv}<.-.(2)>
-      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
-      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
-      \draw (18,4.75)  edge[->,bend left] (copy);
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-      \draw (13,1.25) edge[->,red,thick,bend left] (ret1);
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (12.9,3) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
-      \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
-      \draw node[left=0cm of ret2] { ret-ptr (1) };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=gray!30] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \capbrace[ret3]{(5.1,5)}{(5.1,8)}
-      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-      \draw node[right=0cm of ret3] { ret-ptr (2) };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw[fill=gray!30] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=white] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw (21,7.5) node[devil,opaque=0.5,mirrored,minimum size=1cm] {};
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
-      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
-      \draw (18,8.25)  edge[->,red,thick,bend left] (copy);
-      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw[fill=gray!30] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
-      \draw[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw (9,5.5) node [ellipse, draw, red, thick,minimum height=1.5cm, minimum width=3cm] { };
-      \draw node[red,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker...
-    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
-    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw node[right=0cm of sp1] { stack pointer };
-
-    % stores stack pointer in heap
-    \begin{scope}
-      \fill[gray!20] (14,6) rectangle (18,13);
-      \draw (14,6) -- (14,13);
-      \draw (18,6) -- (18,13);
-    \end{scope}
-    \draw (16,14) node {heap memory};
-    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
-    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
-\end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local caps: Stack clearing prevents Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);
-
-    % attacker calls trusted code again
-    \begin{onlyenv}<+->
-      \inactadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % trusted code clears the stack
-    \begin{onlyenv}<+->
-      \foreach \x in {9,9.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
-    (13,13.5) -- (10,14.5);
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.>
-      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker returns
-    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % trusted code clears stack again
-    \begin{onlyenv}<+->
-      \foreach \x in {7,7.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    % then, the trusted code can return
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 3}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    \begin{scope}
-      \fill[gray!50] (13,0) rectangle (18,13);
-      \draw (13,0) -- (13,13);
-      \draw (18,0) -- (18,13);
-      \draw (15.5,14) node {heap memory};
-    \end{scope}
-
-    \begin{onlyenv}<+->
-      \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \end{onlyenv}
-
-% Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \clip (6.1,2.5) rectangle (8.1,5.5);
-        \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
-      \end{scope}
-      \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
-      \draw node[right=0cm of ret1] { ret-ptr };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<+-.(2)>
-      \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
-    \end{onlyenv}
-
-    \begin{onlyenv}<+>
-      \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
-      \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
-      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \end{onlyenv}
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);
-
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \clip (6.1,5.5) rectangle (8.1,8.5);
-        \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
-      \end{scope}
-      \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
-      \draw node[right=0cm of ret2] { ret-ptr (1) };
-    \end{onlyenv}
-    \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);
-
-    \begin{onlyenv}<+->
-      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
-      \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
-      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \foreach \x in {11.5,12,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \clip (6.1,8.5) rectangle (8.1,11.5);
-        \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
-      \end{scope}
-      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-      \draw node[right=0cm of ret3] { ret-ptr (2) };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
-      \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
-      \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
-      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
-    \end{onlyenv}
-
-    \begin{onlyenv}<+->
-      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
-    \end{onlyenv}
-\end{tikzpicture}
-\end{frame}
-
 \begin{frame}
@@
 
-\end{document}
+\begin{frame}
+  \frametitle{Linear caps prevents attack 3}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+  \stdstackstart
+
+  \begin{scope}
+    \fill[gray!50] (13,0) rectangle (18,13);
+    \draw (13,0) -- (13,13);
+    \draw (18,0) -- (18,13);
+    \draw (15.5,14) node {heap memory};
+  \end{scope}
+
+  \begin{onlyenv}<+->
+    \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+  \end{onlyenv}
+
+  % Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
+  \begin{onlyenv}<+->
+    \begin{scope}
+      \clip (6.1,2.5) rectangle (8.1,5.5);
+      \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
+    \end{scope}
+    \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
+    \draw node[right=0cm of ret1] { ret-ptr };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+  \end{onlyenv}
+
+  \begin{onlyenv}<+-.(2)>
+    \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
+  \end{onlyenv}
+
+  \begin{onlyenv}<+>
+    \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
+    \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+  \end{onlyenv}
+  \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+  \begin{onlyenv}<+->
+    \begin{scope}
+      \clip (6.1,5.5) rectangle (8.1,8.5);
+      \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
+    \end{scope}
+    \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
+    \draw node[right=0cm of ret2] { ret-ptr (1) };
+  \end{onlyenv}
+  \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+    \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \foreach \x in {11.5,12,...,14}
+    {
+      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+    };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \begin{scope}
+      \clip (6.1,8.5) rectangle (8.1,11.5);
+      \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
+    \end{scope}
+    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+    \draw node[right=0cm of ret3] { ret-ptr (2) };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+    \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+    \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+    \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
+  \end{onlyenv}
+\end{tikzpicture}
+
+\end{frame}
+
+%%% Local Variables:
+%%% TeX-master: "presentation-illustrations"
+%%% End:

### 2017-12-13 presentations/illustrations/local-prevention.tex
Restructured document to allow for better compile time during editing.

--- 2017-12-04 presentations/illustrations/presentation-illustrations.tex
@@
-\documentclass{beamer}
-
-\usepackage{ifthen}
-\usepackage{pgf}
-\usepackage{ulem}
-\usepackage{tikz}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{calc}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
-\usepackage{amssymb}
-
-% Active adversary
-\newcommand{\actadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
-
-% Inactive adversary
-\newcommand{\inactadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
-
-% Active stack frame
-\newcommand{\actsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
- }
-}
-
-% Inactive stack frame
-\newcommand{\inactsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }
-}
-
-\newcommand{\capbrace}[3][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
-
-}
-
-\newcommand{\stdstackstart}{
-  \scope
-    \clip (-.1,-.1) rectangle (6.1,15.1);
-    \fill[fill=white] (0,0) rectangle (6,15);
-    \draw (0,0) -- (0,15);
-    \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-  \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-}
-
-% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
-\makeatletter
-\newif\if@showgrid@grid
-\newif\if@showgrid@left
-\newif\if@showgrid@right
-\newif\if@showgrid@below
-\newif\if@showgrid@above
-\tikzset{%
-    every show grid/.style={},
-    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
-    show grid/.default={true},
-    show grid/.cd,
-    labels/.style={font={\sffamily\small},help lines},
-    xlabels/.style={},
-    ylabels/.style={},
-    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
-    true/.style={left,below},
-    false/.style={left=false,right=false,above=false,below=false,grid=false},
-    none/.style={left=false,right=false,above=false,below=false},
-    all/.style={left=true,right=true,above=true,below=true},
-    grid/.is if=@showgrid@grid,
-    left/.is if=@showgrid@left,
-    right/.is if=@showgrid@right,
-    below/.is if=@showgrid@below,
-    above/.is if=@showgrid@above,
-    false,
-}
-
-\def\@showgrid#1{%
-    \begin{scope}[every show grid,show grid/.cd,#1]
-    \if@showgrid@grid
-    \begin{pgfonlayer}{background}
-    \draw [help lines]
-        (current bounding box.south west) grid
-        (current bounding box.north east);
-%
-    \pgfpointxy{1}{1}%
-    \edef\xs{\the\pgf@x}%
-    \edef\ys{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{south west}
-    \edef\xa{\the\pgf@x}%
-    \edef\ya{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{north east}
-    \edef\xb{\the\pgf@x}%
-    \edef\yb{\the\pgf@y}%
-    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
-    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
-    \if@showgrid@below
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
-    }
-    \fi
-    \if@showgrid@above
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
-    }
-    \fi
-    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
-    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
-    \if@showgrid@left
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
-    }
-    \fi
-    \if@showgrid@right
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
-    }
-    \fi
-    \end{pgfonlayer}
-    \fi
-    \end{scope}
-}
-\makeatother
-
-\begin{document}
-\begin{frame}
-  \frametitle{Traditional stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
-    \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
-
-    % traditional:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-
-    % traditional return:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call, attack:
-    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
-
-    % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
-
-    % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
-      skip caller frame};
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+-+(4)>
-      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(3)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    %prepare call: store old return pointer
-    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1); 
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
-      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { new stack pointer };
-    \end{onlyenv}
-
-    % call using capabilities
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker...
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % stores stack pointer in heap
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \fill[gray!20] (14,6) rectangle (18,13);
-        \draw (14,6) -- (14,13);
-        \draw (18,6) -- (18,13);
-      \end{scope}
-      \draw (16,14) node {heap memory};
-      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \end{onlyenv}
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
-    \end{onlyenv}
-
-    % attacker returns
-    \begin{onlyenv}<+>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-
-    % attacker gets called again + uses old stack pointer.
-    \begin{onlyenv}<+>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
-      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+>
-      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<.-.(2)>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
-    \end{onlyenv}
-
-    % attacker calls trusted code again
-%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{onlyenv}<+->
-      \inactadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.-.(1)>
-      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
 \begin{frame}
@@
 \begin{frame}
-    \frametitle{Linear caps: stack pointers}
+  \frametitle{Local Stack Capabilities prevent Attack 3}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
 
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
+    \begin{scope}
+      \fill[gray!50] (13,0) rectangle (18,13);
+      \draw (13,0) -- (13,13);
+      \draw (18,0) -- (18,13);
+      \draw (15.5,14) node {heap memory};
+    \end{scope}
+
+    \begin{onlyenv}<+->
+      \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
     \end{onlyenv}
 
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
+% Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,2.5) rectangle (8.1,5.5);
+        \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
+      \draw node[right=0cm of ret1] { ret-ptr };
     \end{onlyenv}
 
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actsf{(0,5)}{(6,6.5)}{callee stack frame}
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
     \end{onlyenv}
 
-    % Return from callee
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
+    \begin{onlyenv}<+-.(2)>
+      \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
     \end{onlyenv}
 
-    % Caller splices the stack frames
     \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
+      \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
+      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
     \end{onlyenv}
 
+    \begin{onlyenv}<+->
+      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \end{onlyenv}
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,5.5) rectangle (8.1,8.5);
+        \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
+      \draw node[right=0cm of ret2] { ret-ptr (1) };
+    \end{onlyenv}
+    \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \foreach \x in {11.5,12,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,8.5) rectangle (8.1,11.5);
+        \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+      \draw node[right=0cm of ret3] { ret-ptr (2) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+      \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
+    \end{onlyenv}
 \end{tikzpicture}
 \end{frame}
-
-\begin{frame}
-    \frametitle{Linear caps prevent attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-    \draw (16.5,5.5) node {register file};
-    \begin{scope}
-      \clip (15.9,1.1) rectangle (17.1,5.1);
-      \foreach \x in {1,2,...,4}
-      {
-        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
-      };
-    \end{scope}
-    \node (r1) at (16.75,4.5) {};
-    \node (r2) at (16.75,3.5) {};
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,bend left] (sp1);
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<.->
-      \begin{scope}
-        \fill[gray!20] (14,9) rectangle (18,14);
-        \draw (14,9) -- (14,14);
-        \draw (18,9) -- (18,14);
-      \end{scope}
-    \end{onlyenv}
-    \draw<.>[fill=white] (14,10) rectangle node[color=red] {} (18,10.5);
-
-    % Callee stores stack pointer
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (8,5) -- (8,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
-      \draw (16,15) node {heap memory};
-    \end{onlyenv}
-    \begin{onlyenv}<.>
-      \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
-      \draw node[right=0cm of usp] { stack pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-    \end{onlyenv}
-
-    % Return from callee
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
-      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of sp2] { private stack pointer };
-      \draw node[right=0cm of usp] { stack pointer };
-    \end{onlyenv}
-
-    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
-    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
-    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
-  \end{tikzpicture}  
-\end{frame}
-
-\begin{frame}
-  \frametitle{Linear caps prevents attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<+-.(1)>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    % At this point we have a linear point stored in a part of memory that it governs.
-In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
-%    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
-
-    \begin{onlyenv}<+->
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50,right=0cm of sp2] { return data pointer };
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \begin{scope}
-        \fill[fill=gray!50] (0,2.5) rectangle (6,15) node[pos=.5,color=black] {};
-        \draw (0,0) -- (0,15);
-        \draw (6,0) -- (6,15);
-      \end{scope}
-      \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \inactadv{(0,5)}{(6,8)}{callee stack frame}
-      \actsf{(0,8)}{(6,10.5)}{callee (2) stack frame} 
-  \end{onlyenv}
-
-  \draw<+>[teal,very thick] (-0.5,7.5) -- (6.5,11);
-  \draw<.>[teal,very thick] (-0.5,11) -- (6.5,7.5);
-  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { No stack pointer\\for callee (2)!};
-
-  \end{tikzpicture}
-\end{frame}
-
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "presentation-illustrations"
+%%% End:

### 2017-12-13 presentations/illustrations/presentation-illustrations.tex
Restructured document to allow for better compile time during editing.

--- 2017-12-04 presentations/illustrations/presentation-illustrations.tex
@@
 \documentclass{beamer}
+\input{preamble}
+\begin{document}
 
-\usepackage{ifthen}
-\usepackage{pgf}
-\usepackage{ulem}
-\usepackage{tikz}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{calc}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
-\usepackage{amssymb}
+\input{traditional}
 
-% Active adversary
-\newcommand{\actadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
+\input{attacks}
 
-% Inactive adversary
-\newcommand{\inactadv}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \onlyenv<#1>
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
-    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-    \endonlyenv
-  }
-}
+\input{local-prevention}
 
-% Active stack frame
-\newcommand{\actsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
- }
-}
-
-% Inactive stack frame
-\newcommand{\inactsf}[4][]{
-  \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }{
-    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
-  }
-}
-
-\newcommand{\capbrace}[3][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
-
-}
-
-\newcommand{\stdstackstart}{
-  \scope
-    \clip (-.1,-.1) rectangle (6.1,15.1);
-    \fill[fill=white] (0,0) rectangle (6,15);
-    \draw (0,0) -- (0,15);
-    \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-  \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-}
-
-% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
-\makeatletter
-\newif\if@showgrid@grid
-\newif\if@showgrid@left
-\newif\if@showgrid@right
-\newif\if@showgrid@below
-\newif\if@showgrid@above
-\tikzset{%
-    every show grid/.style={},
-    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
-    show grid/.default={true},
-    show grid/.cd,
-    labels/.style={font={\sffamily\small},help lines},
-    xlabels/.style={},
-    ylabels/.style={},
-    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
-    true/.style={left,below},
-    false/.style={left=false,right=false,above=false,below=false,grid=false},
-    none/.style={left=false,right=false,above=false,below=false},
-    all/.style={left=true,right=true,above=true,below=true},
-    grid/.is if=@showgrid@grid,
-    left/.is if=@showgrid@left,
-    right/.is if=@showgrid@right,
-    below/.is if=@showgrid@below,
-    above/.is if=@showgrid@above,
-    false,
-}
-
-\def\@showgrid#1{%
-    \begin{scope}[every show grid,show grid/.cd,#1]
-    \if@showgrid@grid
-    \begin{pgfonlayer}{background}
-    \draw [help lines]
-        (current bounding box.south west) grid
-        (current bounding box.north east);
-%
-    \pgfpointxy{1}{1}%
-    \edef\xs{\the\pgf@x}%
-    \edef\ys{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{south west}
-    \edef\xa{\the\pgf@x}%
-    \edef\ya{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{north east}
-    \edef\xb{\the\pgf@x}%
-    \edef\yb{\the\pgf@y}%
-    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
-    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
-    \if@showgrid@below
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
-    }
-    \fi
-    \if@showgrid@above
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
-    }
-    \fi
-    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
-    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
-    \if@showgrid@left
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
-    }
-    \fi
-    \if@showgrid@right
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
-    }
-    \fi
-    \end{pgfonlayer}
-    \fi
-    \end{scope}
-}
-\makeatother
-
-\begin{document}
-\begin{frame}
-  \frametitle{Traditional stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
-    \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
-
-    % traditional:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-
-    % traditional return:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call, attack:
-    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
-
-    % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
-
-    % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
-      skip caller frame};
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+-+(4)>
-      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(3)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    %prepare call: store old return pointer
-    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1); 
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
-      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { new stack pointer };
-    \end{onlyenv}
-
-    % call using capabilities
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker...
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % stores stack pointer in heap
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \fill[gray!20] (14,6) rectangle (18,13);
-        \draw (14,6) -- (14,13);
-        \draw (18,6) -- (18,13);
-      \end{scope}
-      \draw (16,14) node {heap memory};
-      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \end{onlyenv}
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
-    \end{onlyenv}
-
-    % attacker returns
-    \begin{onlyenv}<+>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-
-    % attacker gets called again + uses old stack pointer.
-    \begin{onlyenv}<+>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
-      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+>
-      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<.-.(2)>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
-    \end{onlyenv}
-
-    % attacker calls trusted code again
-%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{onlyenv}<+->
-      \inactadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.-.(1)>
-      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker...
-    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
-    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw node[right=0cm of sp1] { stack pointer };
-
-    % stores stack pointer in heap
-    \begin{scope}
-      \fill[gray!20] (14,6) rectangle (18,13);
-      \draw (14,6) -- (14,13);
-      \draw (18,6) -- (18,13);
-    \end{scope}
-    \draw (16,14) node {heap memory};
-    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
-    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
-\end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local caps: Stack clearing prevents Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);
-
-    % attacker calls trusted code again
-    \begin{onlyenv}<+->
-      \inactadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % trusted code clears the stack
-    \begin{onlyenv}<+->
-      \foreach \x in {9,9.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
-    (13,13.5) -- (10,14.5);
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.>
-      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker returns
-    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % trusted code clears stack again
-    \begin{onlyenv}<+->
-      \foreach \x in {7,7.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    % then, the trusted code can return
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Linear caps: stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actsf{(0,5)}{(6,6.5)}{callee stack frame}
-    \end{onlyenv}
-
-    % Return from callee
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Caller splices the stack frames
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Linear caps prevent attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-    \draw (16.5,5.5) node {register file};
-    \begin{scope}
-      \clip (15.9,1.1) rectangle (17.1,5.1);
-      \foreach \x in {1,2,...,4}
-      {
-        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
-      };
-    \end{scope}
-    \node (r1) at (16.75,4.5) {};
-    \node (r2) at (16.75,3.5) {};
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,bend left] (sp1);
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<.->
-      \begin{scope}
-        \fill[gray!20] (14,9) rectangle (18,14);
-        \draw (14,9) -- (14,14);
-        \draw (18,9) -- (18,14);
-      \end{scope}
-    \end{onlyenv}
-    \draw<.>[fill=white] (14,10) rectangle node[color=red] {} (18,10.5);
-
-    % Callee stores stack pointer
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (8,5) -- (8,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
-      \draw (16,15) node {heap memory};
-    \end{onlyenv}
-    \begin{onlyenv}<.>
-      \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
-      \draw node[right=0cm of usp] { stack pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-    \end{onlyenv}
-
-    % Return from callee
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
-      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of sp2] { private stack pointer };
-      \draw node[right=0cm of usp] { stack pointer };
-    \end{onlyenv}
-
-    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
-    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
-    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
-  \end{tikzpicture}  
-\end{frame}
-
-\begin{frame}
-  \frametitle{Linear caps prevents attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<+-.(1)>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    % At this point we have a linear point stored in a part of memory that it governs.
-In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
-%    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
-
-    \begin{onlyenv}<+->
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50,right=0cm of sp2] { return data pointer };
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \begin{scope}
-        \fill[fill=gray!50] (0,2.5) rectangle (6,15) node[pos=.5,color=black] {};
-        \draw (0,0) -- (0,15);
-        \draw (6,0) -- (6,15);
-      \end{scope}
-      \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \inactadv{(0,5)}{(6,8)}{callee stack frame}
-      \actsf{(0,8)}{(6,10.5)}{callee (2) stack frame} 
-  \end{onlyenv}
-
-  \draw<+>[teal,very thick] (-0.5,7.5) -- (6.5,11);
-  \draw<.>[teal,very thick] (-0.5,11) -- (6.5,7.5);
-  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { No stack pointer\\for callee (2)!};
-
-  \end{tikzpicture}
-\end{frame}
+\input{linear-prevention}
 

### 2017-12-13 presentations/illustrations/traditional.tex
Restructured document to allow for better compile time during editing.

--- 2017-11-30 presentations/presentation-illustrations.tex
@@
-\documentclass{beamer}
-
-\usepackage{pgf}
-\usepackage{ulem}
-\usepackage{tikz}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
-
-% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
-\makeatletter
-\newif\if@showgrid@grid
-\newif\if@showgrid@left
-\newif\if@showgrid@right
-\newif\if@showgrid@below
-\newif\if@showgrid@above
-\tikzset{%
-    every show grid/.style={},
-    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
-    show grid/.default={true},
-    show grid/.cd,
-    labels/.style={font={\sffamily\small},help lines},
-    xlabels/.style={},
-    ylabels/.style={},
-    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
-    true/.style={left,below},
-    false/.style={left=false,right=false,above=false,below=false,grid=false},
-    none/.style={left=false,right=false,above=false,below=false},
-    all/.style={left=true,right=true,above=true,below=true},
-    grid/.is if=@showgrid@grid,
-    left/.is if=@showgrid@left,
-    right/.is if=@showgrid@right,
-    below/.is if=@showgrid@below,
-    above/.is if=@showgrid@above,
-    false,
-}
-
-\def\@showgrid#1{%
-    \begin{scope}[every show grid,show grid/.cd,#1]
-    \if@showgrid@grid
-    \begin{pgfonlayer}{background}
-    \draw [help lines]
-        (current bounding box.south west) grid
-        (current bounding box.north east);
-%
-    \pgfpointxy{1}{1}%
-    \edef\xs{\the\pgf@x}%
-    \edef\ys{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{south west}
-    \edef\xa{\the\pgf@x}%
-    \edef\ya{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{north east}
-    \edef\xb{\the\pgf@x}%
-    \edef\yb{\the\pgf@y}%
-    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
-    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
-    \if@showgrid@below
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
-    }
-    \fi
-    \if@showgrid@above
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
-    }
-    \fi
-    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
-    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
-    \if@showgrid@left
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
-    }
-    \fi
-    \if@showgrid@right
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
-    }
-    \fi
-    \end{pgfonlayer}
-    \fi
-    \end{scope}
-}
-\makeatother
-
-\begin{document}
 \begin{frame}
@@
     % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \stdstackstart
 
     % first step: capabilities
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,5.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-
-    % call using capabilities
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (6.1,15);
-        \draw[fill=white] (0,2.5) rectangle (6,15.1);
-      \end{scope}
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % attacker...
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+    \begin{onlyenv}<+-+(4)>
+      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
     \end{onlyenv}
-
-    % stores stack pointer in heap
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \clip (13.9,6) rectangle (18.1,13);
-        \draw[fill=gray!20] (14,5.9) rectangle (18,13.1);
-      \end{scope}
-      \draw (16,14) node {heap memory};
-      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \end{onlyenv}
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
-    \end{onlyenv}
-
-    % attacker returns
-    \begin{onlyenv}<+>
+    \begin{onlyenv}<.-.(3)>
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
       \draw node[right=0cm of sp1] { stack pointer };
-      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
     \end{onlyenv}
 
-    % attacker gets called again + uses old stack pointer.
-    \begin{onlyenv}<+>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,12) node[devil,mirrored,minimum size=1cm] {};
-      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+    % prepare call: store old return pointer
+    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
+    \begin{onlyenv}<+-+(2)>
+      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
+      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1); 
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
+      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
+      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
+    \end{onlyenv}
+    \begin{onlyenv}<+->
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { new stack pointer };
+    \end{onlyenv}
+
+    % call using capabilities
+    \begin{onlyenv}<+->
+      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
     \end{onlyenv}
@@
 \end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \draw[fill=white] (0,4) rectangle (6,16);
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % first step: capabilities
-    \begin{onlyenv}<+>
-      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<.-.(2)>
-      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
-    \end{onlyenv}
-
-    % attacker calls trusted code again
-    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.-.(1)>
-      \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
-      \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-
-    % stores stack pointer in heap
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-
-    \draw<.>[green,thick] (13,7.5) -- (19,9);
-    \draw<.>[green,very thick] (19,7.5) -- (13,9);
-    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Clearing prevent Attack 2}
-  TODO
-\end{frame}
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "presentation-illustrations"
+%%% End:

### 2017-12-13 presentations/illustrations/linear-prevention.tex
Added prevention for attack 3.

@@
 
+\begin{frame}
+  \frametitle{Linear caps prevents attack 3}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+  \scope
+    \clip (-.1,-.1) rectangle (6.1,15.1);
+    \fill[fill=white] (0,0) rectangle (6,15);
+    \draw (0,0) -- (0,15);
+    \draw (6,0) -- (6,15);
+    \draw[fill=gray!50] (0,-.5) rectangle (6,1.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+  \endscope
+  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+  \begin{scope}
+    \fill[gray!50] (13,0) rectangle (18,13);
+    \draw (13,0) -- (13,13);
+    \draw (18,0) -- (18,13);
+    \draw (15.5,14) node {heap memory};
+  \end{scope}
+
+\note{Caller calls callee (split stack pointer to make ret-ptr-d)}
+  \begin{onlyenv}<+->
+    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \capbrace[sp]{(6.1,1.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \capbrace[ret1]{(6.1,1.5)}{(6.1,4.5)}
+    \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr-d} };
+    \draw node[right=0cm of ret1] { ret-ptr-d };
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \draw [decorate,decoration={brace,aspect=0.3,amplitude=4pt,raise=4pt},yshift=0pt]
+    (12.9,1) -- (12.9,3) node[draw=black] (ret1c) [black,pos=0.3,xshift=-0.6cm] {};
+    \draw node[fill=gray!50,left=0cm of ret1c] { \phantom{ret-ptr-c} };
+    \draw node[left=0cm of ret1c] { ret-ptr-c };
+  \end{onlyenv}
+
+  \begin{onlyenv}<.-.(2)>
+    \capbrace[sp]{(6.1,4.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+  \end{onlyenv}
+
+\note{Callee saves ret-ptr-d on the heap.}
+  \begin{onlyenv}<+->
+    \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
+  \end{onlyenv}
+\note{Callee saves ret-ptr-c on the heap.}
+  \begin{onlyenv}<+->
+    \draw (13,4.75) edge[red,bend right,thick,->] (ret1c);
+  \end{onlyenv}
+
+\note{Callee calls callee (1)}
+\note{Before the callee jumps, it is assumed that it has checked the global "max height" of its stack pointer!}
+  \begin{onlyenv}<+->
+    \capbrace[ret2]{(6.1,4.5)}{(6.1,7.5)}
+    \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr-d (1)} };
+    \draw node[right=0cm of ret2] { ret-ptr-d (1) };
+  \end{onlyenv}
+  \begin{onlyenv}<.-.(1)>
+    \capbrace[sp]{(6.1,7.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \actsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+  \end{onlyenv}
+
+\note{Callee (1) calls callee (2)}
+  \begin{onlyenv}<+->
+    \capbrace[ret3]{(6.1,7.5)}{(6.1,10.5)}
+    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+    \draw node[right=0cm of ret3] { ret-ptr (2) };
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \capbrace[sp]{(6.1,10.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \inactsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
+
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+    \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+  \end{onlyenv}
+  \begin{onlyenv}<.-.(1)>
+    \actadv{(0,10.5)}{(6,13.5)}{callee stack frame (2)}
+  \end{onlyenv}
+
+\note{Callee (2) has a capability for Callee and can thus access the stored ret-ptr-d}
+  \begin{onlyenv}<+->
+    \draw [decorate,decoration={brace,amplitude=4pt,mirror,raise=4pt},yshift=0pt]
+    (18.1,4) -- (18.1,6) node[draw=black] (calleecap) [black,midway,xshift=0.6cm] {};
+    \draw (18,11) edge[red,bend left,thick,->] (calleecap);
+  \end{onlyenv}
+
+\note{Callee (2) uses ret-ptr-d to return}
+  \begin{onlyenv}<+->
+    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \draw node[fill=white,right=0cm of ret1] { ret-ptr-d };
+    \draw node[fill=white,left=0cm of ret1c] { ret-ptr-c };
+    \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
+  \end{onlyenv}
+\note{Splicing: fails because ends of capabilities don't meet. callee stack frame (1) cannot be empty (it will at least contain 42).}
+  \begin{onlyenv}<+->
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (8,7.5) {Caller tries to splice sp \\and ret-ptr-d, but fails!};  
+  \end{onlyenv}
+\note{Before the caller makes another call/return, it will check that the stack height matches the global max stack height.}
+\end{tikzpicture}
+
+\end{frame}
+
 %%% Local Variables:

### 2017-12-13 presentations/illustrations/attacks.tex
Added comments for attack 3 and the local cap prevention of attack 3. Rectified an omission in the prevetion of attack 3 (forgot first stack clear - not that essential, but it shows how much stack clearing is going to happen)

@@
 
+\note{Caller calls callee}
     \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
@@
 
+\note{Callee has access to some place to store things.}
     \begin{onlyenv}<+->
@@
 
+\note{Callee stores the return pointer they got from caller.}
     \begin{onlyenv}<+->
@@
 
+\note{Callee calls callee (1)}
     \begin{onlyenv}<+->
@@
 
+\note{Callee (1) calls callee (2)}
     \begin{onlyenv}<+->
@@
     \end{onlyenv}
-
+\note{Callee (2) colludes with callee}
     \begin{onlyenv}<+->
@@
     \end{onlyenv}
-
+\note{Callee (2) has access to ret-ptr (the return pointer caller passed to callee)}
     \begin{onlyenv}<+->
@@
 
+\note{Callee (2) uses ret-ptr to return}
     \begin{onlyenv}<+->
@@
 
+\note{This breaks well-bracketedness as ret-ptr (2) and ret-ptr (1) was not returned to}
     \begin{onlyenv}<+->

### 2017-12-13 presentations/illustrations/local-prevention.tex
Added comments for attack 3 and the local cap prevention of attack 3. Rectified an omission in the prevetion of attack 3 (forgot first stack clear - not that essential, but it shows how much stack clearing is going to happen)

@@
 
-% Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
+\note{Caller calls callee}
+\note{Caller clears the stack}
+    \begin{onlyenv}<+->
+      \foreach \x in {5.5,6,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+
+\note{Return pointers always point to the stack as we need to be able to regain the stack pointer itself}
+\note{note: the open ended braces are assumed to stretch over the rest of the stack (in the direction it is open)}
     \begin{onlyenv}<+->
@@
 
+\note{Callee cannot store ret-ptr on the heap as it is local!}
     \begin{onlyenv}<+-.(2)>
@@
 
+\note{Callee tries to store ret-ptr on the stack (just like for attack 2)}
     \begin{onlyenv}<+->
@@
       \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-
     \end{onlyenv}
 
+\note{Callee (1) clears the stack erasing the copy of the ret-ptr!}
     \begin{onlyenv}<+->
@@
 
+\note{The copy of ret-ptr callee (2) hoped to use has been erased}
     \begin{onlyenv}<+->

### 2017-12-14 presentations/illustrations/linear-prevention.tex
added missing capability and removed slides that had been duplicated

@@
     % recurrent parts
-  \stdstackstart
-
-  \begin{scope}
-    \fill[gray!50] (13,0) rectangle (18,13);
-    \draw (13,0) -- (13,13);
-    \draw (18,0) -- (18,13);
-    \draw (15.5,14) node {heap memory};
-  \end{scope}
-
-  \begin{onlyenv}<+->
-    \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-  \end{onlyenv}
-
-  % Note: the return pointer actually spans all of the available stack, but for space reasons, we only draw part of it here.
-  \begin{onlyenv}<+->
-    \begin{scope}
-      \clip (6.1,2.5) rectangle (8.1,5.5);
-      \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
-    \end{scope}
-    \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
-    \draw node[right=0cm of ret1] { ret-ptr };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-    \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
-  \end{onlyenv}
-
-  \begin{onlyenv}<+-.(2)>
-    \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
-  \end{onlyenv}
-
-  \begin{onlyenv}<+>
-    \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
-    \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-  \end{onlyenv}
-  \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);
-
-  \begin{onlyenv}<+->
-    \begin{scope}
-      \clip (6.1,5.5) rectangle (8.1,8.5);
-      \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
-    \end{scope}
-    \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
-    \draw node[right=0cm of ret2] { ret-ptr (1) };
-  \end{onlyenv}
-  \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);
-
-  \begin{onlyenv}<+->
-    \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-    \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
-    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-    \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \foreach \x in {11.5,12,...,14}
-    {
-      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-    };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \begin{scope}
-      \clip (6.1,8.5) rectangle (8.1,11.5);
-      \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
-    \end{scope}
-    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-    \draw node[right=0cm of ret3] { ret-ptr (2) };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-    \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
-    \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
-    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-    \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-    \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
-  \end{onlyenv}
-\end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Linear caps prevents attack 3}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
   \scope
@@
 
-  \begin{onlyenv}<.-.(2)>
+  \begin{onlyenv}<.-.(3)>
     \capbrace[sp]{(6.1,4.5)}{(6.1,15)}

### 2017-12-15 fully-abstract-comp/technical-report.tex
fix typo, did not compile
typos
removed the 2nd part of the K-relation
small updates
Removed linearity from seals - it seemed to add complexity without being useful in any sensible way.

@@
 % \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
-Source language domains:
 \[
@@
     &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
-    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\linbnf,\sealbasebnf,\sealendbnf,\sigma}\\
+    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
@@
   \[
-    \range{\seal{\_,\sigma_\baddr,\sigma_\eaddr,\_}} = [\sigma_\baddr,\sigma_\eaddr]
+    \range{\seal{\sigma_\baddr,\sigma_\eaddr,\_}} = [\sigma_\baddr,\sigma_\eaddr]
   \]
@@
       \cb = ((\_,\linear),\_,\_,\_) \tor\\
-      \cb = \seal{\linear,\_,\_,\_} \tor\\
       \cb = \sealed{\_,\cb'} \tand \isLinear{\cb'} 
@@
   \sem{\tfail}(\Phi) = & \; \failed \\
-  \sem{\thalt}(\Phi) = & \; \haltde
+  \sem{\thalt}(\Phi) = & \; \halted
 \end{align*}
@@
                                        \Phi(r_2) = ((\_,\_),\_,\_,\aaddr) \\
-                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\_,\aaddr} \\
+                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\aaddr} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\aaddr} } \\
@@
                                        \Phi(r_2) = ((\_,\_),\baddr,\_,\_) \\
-                                       \quad \tor \Phi(r_2) = \seal{\_,\baddr,\_,\_} \\
+                                       \quad \tor \Phi(r_2) = \seal{\baddr,\_,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\baddr,\_,\_} } \\
@@
                                        \Phi(r_2) = ((\_,\_),\_,\eaddr,\_) \\
-                                       \quad \tor \Phi(r_2) = \seal{\_,\_,\eaddr,\_} \\
+                                       \quad \tor \Phi(r_2) = \seal{\_,\eaddr,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\eaddr,\_} } \\
@@
                                        \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                       \quad \tor \Phi(r_2) = \seal{\lin,\_,\_,\_} \\
                                        \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} \tand l = \linear}\\
                                      \end{array} \\
+                                     \updPcAddr{\Phi\updReg{r_1}{\normal}} & \Phi(r_2) = \seal{\_,\_,\_}\\
                                      \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
@@
                                       \Phi(\rn) = n \tand \\
-                                      \Phi(r) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
-                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma + n}
+                                      \Phi(r) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
+                                      c = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma + n}
                                     \end{array}
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
+                                      \Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}\\
-                                      c = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
+                                      c = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
                                     \end{array} \\
@@
                                       \Phi(r_2) \in \SealableCaps \tand \\
-                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
+                                      \Phi(r_3) = \seal{\sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
                                       \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
@@
                                  \begin{array}[t]{l}
-                                   \Phi(r_3) = \seal{\lin,\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
+                                   \Phi(r_3) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
                                    \Phi(\rn_4) = n \tand\\
                                    \sigma_\baddr \leq n \tand n < \sigma_\eaddr \tand \\
-                                   c_1 = \seal{\lin,\sigma_\baddr,n,\sigma} \tand \\
-                                   c_2 = \seal{\lin,n+1,\sigma_\eaddr,\sigma} \tand \\
+                                   c_1 = \seal{\sigma_\baddr,n,\sigma} \tand \\
+                                   c_2 = \seal{n+1,\sigma_\eaddr,\sigma} \tand \\
                                    w = \linCons{\Phi(r_3)}
@@
                                 \begin{array}[t]{l}
-                                  \Phi(r_2) = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,2},\_} \tand \\
-                                  \Phi(r_3) = \seal{\lin,\sigma_{\baddr,3},\sigma_{\eaddr,3},\sigma} \tand \\
+                                  \Phi(r_2) = \seal{\sigma_{\baddr,2},\sigma_{\eaddr,2},\_} \tand \\
+                                  \Phi(r_3) = \seal{\sigma_{\baddr,3},\sigma_{\eaddr,3},\sigma} \tand \\
                                   \sigma_{\eaddr,2}+1 = \sigma_{\baddr,3} \tand \sigma_{\baddr,2} \leq \sigma_{\eaddr,2} \tand\\
                                   \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
-                                  c = \seal{\lin,\sigma_{\baddr,2},\sigma_{\eaddr,3},\sigma} \tand \\
+                                  c = \seal{\sigma_{\baddr,2},\sigma_{\eaddr,3},\sigma} \tand \\
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
@@
                                                    \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
-                                                   \mem(\baddr) = \seal{\_,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
+                                                   \mem(\baddr) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
                                                    \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
@@
   \item $w \in \ints$
-  \item $w = \seal{\_,\_,\_,\_}$
+  \item $w = \seal{\_,\_,\_}$
   \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
@@
   \item $\ms$ respects write-xor-execute.
-  \item $\ms(\baddr_\code) = \seal{\normal,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ where these are somehow the sufficient amount of seals.
+  \item $\ms(\baddr_\code) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ where these are somehow the sufficient amount of seals.
   \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
@@
       \quad \forall W_\lrr, W_\lrm \ldotp  \\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk::(\opc,\ms_{\stk,f}),\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,(\opc,\ms_{\stk,f}) :: \stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
       \qquad\opc = c_c \wedge\\
       \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
-      \qquad\npair[n']{\array{l}\src{\reg_S\update{\rstk}{\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}}},\\
-                               {\reg_T\update{\rstk}{((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)}}\endarray}\in \lrr(\{r_\stk\})(W_\lrr) \wedge\\ 
+      \qquad\npair[n']{\left(\array{l}\src{\reg_S\update{\rstk}{\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}}},\\
+                               {\reg_T\update{\rstk}{((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)}}\endarray \right)}\in \lrr(\{r_\stk\})(W_\lrr) \wedge\\ 
       \qquad\npair[n']{[(\opc,\opc)]} \in \lrp(W')\\
@@
     \end{array}
-  \right\}\\ \cup
-    & \left\{ \npair{\left( k_{c,S},k_{d,S},k_{c,T},k_{d,T} \right)} \middle|
-      \begin{array}{l}
-        \pi_\sigma(k_{c,S}) = \pi_\sigma(k_{d,S}) \wedge \pi_\sigma(k_{c,T}) = \pi_\sigma(k_{d,T}) \wedge \npair{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \\
-        \forall n' \leq n, W' \pubft W \ldotp \\
-        \quad \npair[n']{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \Rightarrow\\
-        \qquad \forall \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W'_\lrr) \ldotp 
-               \forall \memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W'_\lrm} \ldotp\\
-        \qquad \quad \npair[n']{\left(
-        \arraycolsep=0pt\array{l}
-          (\src{\ms_S},\src{\reg_S}\update{\pcreg,\rdata}{\pi_\vsc(\src{k_{c,S}}),\pi_\vsc(\src{k_{d,S}})},\src{\stk},\src{\ms_\stk})\\
-          (\ms_T,\reg_T\update{\pcreg,\rdata}{\pi_\vsc(k_{c,T}),\pi_\vsc(k_{d,T})})
-        \endarray\right)} \in \lro(W' \oplus W'_\lrr \oplus W'_\lrm)
-      \end{array}
   \right\}
+% \\ \cup
+%     & \left\{ \npair{\left( k_{c,S},k_{d,S},k_{c,T},k_{d,T} \right)} \middle|
+%       \begin{array}{l}
+%         \pi_\sigma(k_{c,S}) = \pi_\sigma(k_{d,S}) \wedge \pi_\sigma(k_{c,T}) = \pi_\sigma(k_{d,T}) \wedge \npair{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \\
+%         \forall n' \leq n, W' \pubft W \ldotp \\
+%         \quad \npair[n']{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \Rightarrow\\
+%         \qquad \forall \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W'_\lrr) \ldotp 
+%                \forall \memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W'_\lrm} \ldotp\\
+%         \qquad \quad \npair[n']{\left(
+%         \arraycolsep=0pt\array{l}
+%           (\src{\ms_S},\src{\reg_S}\update{\pcreg,\rdata}{\pi_\vsc(\src{k_{c,S}}),\pi_\vsc(\src{k_{d,S}})},\src{\stk},\src{\ms_\stk}),\\
+%           (\ms_T,\reg_T\update{\pcreg,\rdata}{\pi_\vsc(k_{c,T}),\pi_\vsc(k_{d,T})})
+%         \endarray\right)} \in \lro(W' \oplus W'_\lrr \oplus W'_\lrm)
+%       \end{array}
+%      \right\}
   \end{array}
@@
     \begin{array}{l}
-      \vsc_c = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge \\
+      \vsc_C = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge \\
       \forall (\src{c},c') \in H_\sigma(\sigma) \ldotp \\
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\lin,\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\lin,\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
     \right\} \cup \\
@@
 \]
-\lau{28-11-2017: Do we want the execute condition for $\rwx$ is we are doing write xor execute?}
+\lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}
 

### 2017-12-18 fully-abstract-comp/technical-report.tex
A bunch of small corrections for the first part of the document (up until operational semantics) removed some things that we no longer assume

@@
 \newcommand{\rdata}{\mathrm{r}_\mathrm{data}}
+\newcommand{\rtmp}[1]{\mathrm{r}_\mathrm{t#1}}
 
@@
     &\cbnf & \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
-    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
+    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
     &\Reg & \defeq & \RegName \fun \Word\\
@@
     &\MemSeg & \defeq & \Addr \parfun \Word \\
-    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg \times \Addr}\\
+    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg}\\
     \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
@@
 \]
-\lau{06-09-2017: Should seals also have the option to be linear?}
-\lau{13-09-2017: For now I have added linearity to the seals.}
+%\lau{06-09-2017: Should seals also have the option to be linear?}
+%\lau{13-09-2017: For now I have added linearity to the seals.}
+%\lau{15-12-2017: Removed linearity from seal - I could not come up with a single good reason to have it as we have them here.}
 \dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
@@
 In the source language, $\src{\Stack}$ is a call stack that contains the data for each call.
-The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc, 2) caller's private stack, and 3) the address of the old stack pointer.
+The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc and 2) caller's private stack.
+\lau{18-12-2017: Stack frame used to contain the address of the old stack pointer. This should be inferable from other information, so it was removed.}
 
@@
 In the target language, this will be the base address of some linear capability.
-The base address can be any address on the machine, but it will have to remain the same.
+The base address can be any address on the machine, but it will have to remain the same during all of execution.
 We write this constant as
@@
 \]
-
-More convenient definitions
-\[
-  \begin{array}{rcl}
-    \Phi(r) & \defeq & \Phi.\reg(r)
-  \end{array}
-\]
-where $r\in \RegName$.
 
@@
 
-\dominique{4-12-2017: some instructions need to be able to take an integer rather than a natural, e.g. cca (as used in the implementation of call).}
+%Lau: resolved:
+%\dominique{4-12-2017: some instructions need to be able to take an integer rather than a natural, e.g. cca (as used in the implementation of call).}
 
 The source machine is also a capability machine with memory capabilities and sealed capabilities.
-Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities.
+Unlike the target machine, the source machine has a built in stack along with special stack and return tokens used in place of the actual capabilities.
 The syntax of the source machine language is as follows:
@@
   \begin{array}{rcl}
-    n & \in & \nats \\
     o & \in & \nats \\
-    \src{r} &\in &  \RegName\\
-    \src{\rn} &\defbnf & \src{r} \mid n \\
-    \src{i} & \defbnf &  \trg{i} \mid \scall{o}{\src{r}}{\src{r}}
+    \src{i} & \defbnf &  \trg{i} \mid \scall{o}{r}{r}
   \end{array}
@@
 The source machine allows for variable length instructions which we utilize for $\scall{o}{}{}$.
-In other words, when we $\scall{o}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction.
+In other words, when $\scall{o}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction (and be within the range of the pc capability).
 This is described in more detail when we present the decoding function.
@@
 \subsection{Permissions}
-\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.} % I am not sure why I did not think the opaque capabilities where necessary, but I have a feeling we talked about this in Leuven and came to the conclusion that they were necessary.
+%\lau{23-03-2017: I am not sure we need the opaque capabilities anymore.} % I am not sure why I did not think the opaque capabilities where necessary, but I have a feeling we talked about this in Leuven and came to the conclusion that they were necessary.
 \begin{figure}[!h]
@@
 \end{figure}
+\lau{13-09-2017: Note that most of the discussion belove is now captured in Section~\ref{sec:linear-cap}}
 \lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear.
@@
 This would call for a ``load from offset'' operation to be able to load an in code stored seal (see the figure I drew about linking/seals).}
-\lau{13-09-2017: Note that most of the above discussion is now captured in Section~\ref{sec:linear-cap}}
 
@@
 
+For convenience, we introduce the following notation:
+\[
+  \begin{array}{rcl}
+    \Phi(r) & \defeq & \Phi.\reg(r)
+  \end{array}
+\]
+where $r\in \RegName$.
+
 For $\rn$ which can be a register or an integer, we take

### 2017-12-18 presentations/illustrations/preamble.tex
preamble

--- 2017-11-28 presentations/presentation-illustrations.tex
@@
-\documentclass{beamer}
-
+\usepackage{ifthen}
 \usepackage{pgf}
@@
 \usepackage{tikzpeople}
+\usepackage{amssymb}
+
+% Active adversary
+\newcommand{\actadv}[4][]{
+  \ifthenelse{\equal{#1}{}}{
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+    \onlyenv<#1>
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \endonlyenv
+  }
+}
+
+% Inactive adversary
+\newcommand{\inactadv}[4][]{
+  \ifthenelse{\equal{#1}{}}{
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
+    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+    \onlyenv<#1>
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
+    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \endonlyenv
+  }
+}
+
+% Active stack frame
+\newcommand{\actsf}[4][]{
+  \ifthenelse{\equal{#1}{}}{
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
+ }
+}
+
+% Inactive stack frame
+\newcommand{\inactsf}[4][]{
+  \ifthenelse{\equal{#1}{}}{
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }{
+    \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
+    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+  }
+}
+
+\newcommand{\capbrace}[3][sp1]{
+  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
+
+}
+
+\newcommand{\stdstackstart}{
+  \scope
+    \clip (-.1,-.1) rectangle (6.1,15.1);
+    \fill[fill=white] (0,0) rectangle (6,15);
+    \draw (0,0) -- (0,15);
+    \draw (6,0) -- (6,15);
+    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+  \endscope
+  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+}
 
@@
 \makeatother
-
-\begin{document}
-\begin{frame}
-  \frametitle{Traditional stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
-    \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
-
-    % traditional:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-
-    % traditional return:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call, attack:
-    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
-
-    % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \draw<.> (3,7) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
-
-    % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
-      skip caller frame};
-  \end{tikzpicture}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % first step: capabilities
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,5.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-
-    % call using capabilities
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-
-    % stores stack pointer in heap
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-
-    % attacker returns
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<+>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,6.5) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.>[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
-
-    % attacker gets called again + uses old stack pointer.
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<+>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,6.5) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (spold);
-    \draw<.>[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (3,12) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,13.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % first step: capabilities
-    \draw<+>[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,4) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,7) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> (3,5.5) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
-
-    % attacker calls trusted code again
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.>[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);
-
-    % attacker calls trusted code again
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.>[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \draw<.>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,12) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> (3,10.5) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (10.5,4) -- (10.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-    (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (spold);
-    \draw<.> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-
-    % stores stack pointer in heap
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-
-    \draw<.>[green,thick] (13,7.5) -- (19,9);
-    \draw<.>[green,very thick] (19,7.5) -- (13,9);
-    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Clearing prevent Attack 2}
-  TODO
-\end{frame}
-\end{document}

### 2017-12-18 fully-abstract-comp/technical-report.tex
filled in some blanks

@@
 \newcommand{\vsc}{\var{sc}}
-\newcommand{\cb}{\var{cb}}
+\newcommand{\cb}{\vsc}
 \newcommand{\baddr}{\var{b}}
@@
   \item This is needed for correctness. See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
-  \end{itemize}
-\item Programs have undefined behavior when passing ret/stk pointers.
+  \item It is only used for the call instruction.
+  \end{itemize}
+% Lau: 18-12-2017: I do not think this is true anymore?
+% \item Programs have undefined behavior when passing ret/stk pointers.
 \end{itemize}
@@
 \paragraph{Step relation}
-The instruction $\scall{o}{}{}$ has length $\calllen$ which means that it does not fit on one memory address.
+The instruction $\scall{o}{}{}$ has length $\calllen$ which means that it does not fit in one memory address.
 In fact, $\scall{o}{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
@@
         \vdots \\
-        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen]{o}{r_1}{r_2}
+        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen-1]{o}{r_1}{r_2}
       \end{array}
@@
                                                     \src{\eaddr < \aaddr + \calllen-1) \tand} \\
-                                                    \baddr \leq \aaddr \leq \eaddr\tand\\
+                                                    \withinBounds{\Phi(\pcreg)} \tand \\
                                                     \exec{\Phi(\pcreg)}
@@
 \label{par:call-impl}
-\lau{16-10-2017: Fill this paragraph with $\scall{o}{}{}$ implementation, i.e., the instructions $\scall[1]{o}{r_1}{r_2} \dots \scall[\calllen]{o}{r_1}{r_2}$ is equal to.
+This paragraph contains the implementation of $\scall{o}{}{}$.
+That is each of the instructions in the implementation corresponds to $\scall[1]{o}{r_1}{r_2} \dots \scall[\calllen-1]{o}{r_1}{r_2}$, respectively.
+
+%\lau{16-10-2017: Fill this paragraph with $\scall{o}{}{}$ implementation, i.e., the instructions $\scall[1]{o}{r_1}{r_2} \dots \scall[\calllen]{o}{r_1}{r_2}$ is equal to.
 Add convenience notation like $\update{\mem.a}{\scall{o}{r_1}{r_2}} = \dots$}
@@
 \begin{itemize}
-\item Store the end address of the stack to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
+\item Store 42 (it could be anything) to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
 \item Get the current address of the stack pointer and split according to it.
@@
 
+For convenience, we will add the convention that the memory update $\update{\mem.a}{\scall{o}{r_1}{r_2}}$ corresponds to
+\[
+  \begin{array}{l}
+    \update{\mem.a}{\scall[0]{o}{r_1}{r_2}}\\
+    \update{\mem.a}{\scall[1]{o}{r_1}{r_2}}\\
+    \dots\\
+    \update{\mem.a}{\scall[\calllen-1]{o}{r_1}{r_2}}\\
+  \end{array}
+\]
+
 \subsubsection{Helpful functions, sets, and conventions}
@@
     \src{\baddr} & \src{\text{if $\vsc = \stkptr{\_,\baddr,\_,\_}$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrd{\_,(\baddr,\_)}$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrc{\_,(\_,\_,\baddr,\_,\_)}$}}
+    \src{\baddr} & \src{\text{if $\vsc = \retptrd(\_,(\baddr,\_))$}} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\baddr,\_,\_))$}}
   \end{cases}
@@
     \src{\eaddr} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrd{\_,(\_,\eaddr)}$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrc{\_,(\_,\_,\_,\eaddr,\_)}$}}
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrd(\_,(\_,\eaddr))$}} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\_,\eaddr,\_))$}}
   \end{cases}
@@
     \src{\linear} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
-    \src{\linear} & \src{\text{if $\vsc = \retptrd{\_,(\_,\_)}$}} \\ 
-    \src{\lin} & \src{\text{if $\vsc = \retptrc{\_,(\_,\lin,\_,\_,\_)}$}}
+    \src{\linear} & \src{\text{if $\vsc = \retptrd(\_,(\_,\_))$}} \\ 
+    \src{\lin} & \src{\text{if $\vsc = \retptrc(\_,(\_,\lin,\_,\_,\_))$}}
   \end{cases}
@@
 \[
-  \updPcAddr{c} \defeq \dots
-\]
-
-\[
-  \readAllowed{} \defeq \dots
-\]
-
-\[
-  \writeAllowed{} \defeq \dots
+  \updPcAddr{\Phi} \defeq 
+  \begin{cases}
+    \Phi\update{\pcreg}{((\perm,\lin),\baddr,\eaddr,\aaddr+1)} & \text{if $\Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$}\\
+    \failed & \totherwise
+  \end{cases}
+\]
+
+\[
+  \readAllowed{} \defeq \{ \rwx,\rw,\rx,\ro \}
+\]
+
+\[
+  \writeAllowed{} \defeq \{ \rwx, \rw \}
 \]
@@
 \[
-  \exec{\cb} \defeq \dots
-\]
-
-\[
-  \nonExec{\cb} \defeq \dots
-\]
-
-\[
-  \withinBounds{c} \defeq \dots
+  \exec{\cb} \defeq 
+      \cb = ((\perm, \_), \_, \_, \_) \tand \perm \in \{\rwx,\rx\} 
+\]
+
+\[
+  \nonExec{\cb} \defeq \neg \exec{\cb}
+\]
+
+\[
+  \withinBounds{\vsc} \defeq 
+  \begin{cases}
+    \baddr \leq \aaddr \leq \eaddr & 
+    \arraycolsep=0pt
+    \begin{array}{l}
+      \vsc = ((\_,\_),\baddr,\eaddr,\aaddr) \src{\tor}\\
+      \src{\vsc = \stkptr{\_,\baddr,\eaddr,\aaddr}}
+    \end{array}\\
+    \sigma_\baddr \leq \sigma_\aaddr \leq \sigma_\eaddr & \vsc = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \\
+    \bot & \totherwise
+  \end{cases}
 \]
@@
                                       \Phi(r_3) = \seal{\sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
-                                      \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
+                                      \withinBounds{\Phi(r_3)} \tand \\
                                       \vsc = \sealed{\sigma,\Phi(r_2)} \tand\\

### 2017-12-18 presentations/illustrations/traditional.tex
Simplify slides a bit: direction of stack is clear, don't show old return data pointer, just new one.

@@
       \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
+      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
     \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
+    \draw (-1.5,0) node {};
+    \draw (-1.5,15) node {};
+    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
 
@@
     \begin{onlyenv}<.-.(1)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+      \draw (7.5,4) node[right,fill=gray!50] (rp1l) {return pointer data};
     \end{onlyenv}
@@
       \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      \draw (0,5.75) edge[->,out=180,in=225] (oldframe1); 
+      % \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      % (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+      % \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
     \end{onlyenv}

### 2017-12-19 presentations/illustrations/traditional.tex
Presentation illustrations: move slides about stack and return capabilities to separate file.

@@
 
-\begin{frame}
-  \frametitle{Stack and return capabilities}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % first step: capabilities
-    \begin{onlyenv}<+-+(4)>
-      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(3)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \begin{onlyenv}<.-.(1)>
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw (7.5,4) node[right,fill=gray!50] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    % prepare call: store old return pointer
-    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \begin{onlyenv}<+-+(2)>
-      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-      % \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      % (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-      % \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
-      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return pointer data};
-    \end{onlyenv}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { new stack pointer };
-    \end{onlyenv}
-
-    % call using capabilities
-    \begin{onlyenv}<+->
-      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \end{onlyenv}
-  \end{tikzpicture}
-\end{frame}
 %%% Local Variables:

### 2017-12-19 presentations/illustrations/attacks.tex
Presentation illustrations: move text slide to main file.
Presentation illustrations: tune attack 3 slide to remove confusion about heap vs code memory.

@@
 \begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{scope}
-    \draw (2.5,11) node {heap memory};
+    \draw (2.5,11) node {code memory};
 
     \begin{scope}
-    \clip (12.9,-.1) rectangle (18.1,11.1);
-    \fill[fill=gray!50] (13,0) rectangle (18,11);
-    \draw (13,0) -- (13,11);
-    \draw (18,0) -- (18,11);
+    \clip (12.9,3.9) rectangle (18.1,11.1);
+    \fill[fill=gray!50] (13,4) rectangle (18,11);
+    \draw (13,4) -- (13,11);
+    \draw (18,4) -- (18,11);
     \end{scope}
-    \draw (15.5,12) node {more heap memory};
+    \draw (15.5,12) node {more code memory};
+
+    \begin{scope}
+    \clip (12.9,-.1) rectangle (18.1,2.1);
+    \fill[fill=gray!50] (13,0) rectangle (18,2);
+    \draw (13,0) -- (13,2);
+    \draw (18,0) -- (18,2);
+    \end{scope}
+    \draw (15.5,3) node {heap memory};
 
@@
       \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=white] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
       \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
@@
       (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
-      \draw (18,4.75)  edge[->,bend left] (copy);
+      \draw (18,5.5)  edge[->,bend left] (copy);
     \end{onlyenv}
@@
     \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (12.9,3) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=3pt,raise=4pt},yshift=0pt]
+      (12.9,5) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
       \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
-      \draw node[left=0cm of ret2] { ret-ptr (1) };
+      \draw node[left=0cm of ret2] (retptr1) { ret-ptr (1) };
     \end{onlyenv}
@@
       \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=gray!30] (13,3) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
       \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
@@
       \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-      \draw node[right=0cm of ret3] { ret-ptr (2) };
+      \draw node[right=0cm of ret3] (retptr2) { ret-ptr (2) };
     \end{onlyenv}
@@
     \begin{onlyenv}<+->
-      \draw (9,5.5) node [ellipse, draw, red, thick,minimum height=1.5cm, minimum width=3cm] { };
+      \draw node [ellipse, fit=(retptr1) (retptr2), draw, red] { };
       \draw node[red,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  

### 2017-12-19 presentations/illustrations/capabilities.tex

--- 2017-11-30 presentations/presentation-illustrations.tex
@@
-\documentclass{beamer}
-
-\usepackage{pgf}
-\usepackage{ulem}
-\usepackage{tikz}
-\usetikzlibrary{backgrounds}
-\usetikzlibrary{positioning}
-\usetikzlibrary{patterns}
-\usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
-
-% stolen from https://tex.stackexchange.com/questions/14225/is-there-the-easiest-way-to-toggle-show-hide-navigational-grids-in-tikz/14230
-\makeatletter
-\newif\if@showgrid@grid
-\newif\if@showgrid@left
-\newif\if@showgrid@right
-\newif\if@showgrid@below
-\newif\if@showgrid@above
-\tikzset{%
-    every show grid/.style={},
-    show grid/.style={execute at end picture={\@showgrid{grid=true,#1}}},%
-    show grid/.default={true},
-    show grid/.cd,
-    labels/.style={font={\sffamily\small},help lines},
-    xlabels/.style={},
-    ylabels/.style={},
-    keep bb/.code={\useasboundingbox (current bounding box.south west) rectangle (current bounding box.north west);},
-    true/.style={left,below},
-    false/.style={left=false,right=false,above=false,below=false,grid=false},
-    none/.style={left=false,right=false,above=false,below=false},
-    all/.style={left=true,right=true,above=true,below=true},
-    grid/.is if=@showgrid@grid,
-    left/.is if=@showgrid@left,
-    right/.is if=@showgrid@right,
-    below/.is if=@showgrid@below,
-    above/.is if=@showgrid@above,
-    false,
-}
-
-\def\@showgrid#1{%
-    \begin{scope}[every show grid,show grid/.cd,#1]
-    \if@showgrid@grid
-    \begin{pgfonlayer}{background}
-    \draw [help lines]
-        (current bounding box.south west) grid
-        (current bounding box.north east);
-%
-    \pgfpointxy{1}{1}%
-    \edef\xs{\the\pgf@x}%
-    \edef\ys{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{south west}
-    \edef\xa{\the\pgf@x}%
-    \edef\ya{\the\pgf@y}%
-    \pgfpointanchor{current bounding box}{north east}
-    \edef\xb{\the\pgf@x}%
-    \edef\yb{\the\pgf@y}%
-    \pgfmathtruncatemacro\xbeg{ceil(\xa/\xs)}
-    \pgfmathtruncatemacro\xend{floor(\xb/\xs)}
-    \if@showgrid@below
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [below,show grid/labels,show grid/xlabels] at (\X,\ya) {\X};
-    }
-    \fi
-    \if@showgrid@above
-    \foreach \X in {\xbeg,...,\xend} {
-        \node [above,show grid/labels,show grid/xlabels] at (\X,\yb) {\X};
-    }
-    \fi
-    \pgfmathtruncatemacro\ybeg{ceil(\ya/\ys)}
-    \pgfmathtruncatemacro\yend{floor(\yb/\ys)}
-    \if@showgrid@left
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [left,show grid/labels,show grid/ylabels] at (\xa,\Y) {\Y};
-    }
-    \fi
-    \if@showgrid@right
-    \foreach \Y in {\ybeg,...,\yend} {
-        \node [right,show grid/labels,show grid/ylabels] at (\xb,\Y) {\Y};
-    }
-    \fi
-    \end{pgfonlayer}
-    \fi
-    \end{scope}
-}
-\makeatother
-
-\begin{document}
-\begin{frame}
-  \frametitle{Traditional stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize higher stack frames...};
-    \end{scope}
-    \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
-
-    % traditional:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-
-    % traditional return:
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.> (8,7) node[draw=black] (sp0) {};
-    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
-
-    % traditional call, attack:
-    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
-
-    % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
-    \draw<.> (3,7) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
-
-    % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
-    \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
-      skip caller frame};
-  \end{tikzpicture}
-
-\end{frame}
-
 \begin{frame}
@@
     % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \stdstackstart
 
     % first step: capabilities
-    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,5.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    \begin{onlyenv}<+-+(4)>
+      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(3)>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+      \draw node[right=0cm of sp1] { stack capability };
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+      \draw (7.5,4) node[right,fill=gray!50] (rp1l) {return capability data};
+    \end{onlyenv}
+
+    % prepare call: store old return capability
+    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\scriptsize old return cap code} (6,6.5);
+    \begin{onlyenv}<+-+(2)>
+      \draw[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return cap data} (6,6);
+      % \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      % (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+      % \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
+      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
+      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return capability data};
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { new stack capability };
+    \end{onlyenv}
 
     % call using capabilities
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-    \draw<.> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
-    \draw<.>[fill=gray!50] (0,5.5) rectangle node {\scriptsize old return pointer data} (6,6);
-    \draw<.>[fill=gray!50] (0,6) rectangle node {\scriptsize old return pointer code} (6,6.5);
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
-      (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
-    \draw<.> (0,5.75) edge[->,out=180,in=225] (oldframe1);
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (6.1,15);
-        \draw[fill=white] (0,2.5) rectangle (6,15.1);
-      \end{scope}
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % attacker...
-    \begin{onlyenv}<+-+(2)>
+    \begin{onlyenv}<+->
       \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
       \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,8) node[devil,mirrored,minimum size=1cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % stores stack pointer in heap
-    \begin{onlyenv}<+->
-      \begin{scope}
-        \clip (13.9,6) rectangle (18.1,13);
-        \draw[fill=gray!20] (14,5.9) rectangle (18,13.1);
-      \end{scope}
-      \draw (16,14) node {heap memory};
-      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \end{onlyenv}
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
-    \end{onlyenv}
-
-    % attacker returns
-    \begin{onlyenv}<+>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
-    \end{onlyenv}
-
-    % attacker gets called again + uses old stack pointer.
-    \begin{onlyenv}<+>
-      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
-      \draw[fill=white] (0,10.5) rectangle (6,13.5) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,12) node[devil,mirrored,minimum size=1cm] {};
-      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
-      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
     \end{onlyenv}
@@
 
-\begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-      \draw[fill=white] (0,4) rectangle (6,16);
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % first step: capabilities
-    \begin{onlyenv}<+>
-      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
-    \end{onlyenv}
-
-    % attacker stores copy of stack pointer high in the stack
-    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
-    \begin{onlyenv}<.-.(2)>
-      \draw[fill=white] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-      \draw (3,5.5) node[devil,mirrored,minimum size=1cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
-    \end{onlyenv}
-
-    % attacker calls trusted code again
-    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % attacker uses old stack pointer
-    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-    \begin{onlyenv}<.-.(1)>
-      \draw[fill=white] (0,9) rectangle (6,12) node[pos=.5] {\footnotesize callee (3) stack frame};
-      \draw (3,10.5) node[devil,mirrored,minimum size=1cm] {};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Capabilities Problems}
-  \begin{itemize}
-  \item Stack capabilities are not a panacea
-  \item Stack pointers must not leave stack frame (through memory or stack)
-  \item Similar problems for return pointers (omitted)
-  \item Solution: local capabilities?
-    \begin{itemize}
-    \item Make stack and return capabilities local
-    \item Cannot leave registers (except on stack) (\sout{Attack 1})
-    \item Stack clearing on boundary crossings (\sout{Attack 2})
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Local Stack Capabilities prevent Attack 1}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15.1);
-      \draw[fill=gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
-      \draw[draw=gray!80] (0,0) -- (0,15);
-      \draw[draw=gray!80] (6,0) -- (6,15);
-      \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-    \end{scope}
-
-    \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-    % attacker...
-    \draw<+>[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
-    \draw<.>[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
-    \begin{scope}
-      \clip (-.1,-.1) rectangle (6.1,15);
-      \draw<.>[fill=white] (0,9.5) rectangle (6,15.1);
-    \end{scope}
-    \draw<.> (3,8) node[devil,mirrored,minimum size=1cm] {};
-    \draw<.> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw<.> node[right=0cm of sp1] { stack pointer };
-
-    % stores stack pointer in heap
-    \begin{scope}
-      \clip (13.9,6) rectangle (18.1,13);
-      \draw<.>[fill=gray!20] (14,5.9) rectangle (18,13.1);
-    \end{scope}
-    \draw<.> (16,14) node {heap memory};
-    \draw<.>[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
-
-    \draw<.>[green,thick] (13,7.5) -- (19,9);
-    \draw<.>[green,very thick] (19,7.5) -- (13,9);
-    \draw node[green,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw] () at (14,4) {Stack pointer is local!};
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Stack Clearing prevent Attack 2}
-  TODO
-\end{frame}
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "presentation-illustrations"
+%%% End:

### 2017-12-19 presentations/illustrations/local-prevention.tex
Clarify attack 3 in local setting.

@@
     \begin{scope}
-      \fill[gray!50] (13,0) rectangle (18,13);
-      \draw (13,0) -- (13,13);
-      \draw (18,0) -- (18,13);
-      \draw (15.5,14) node {heap memory};
+      \fill[gray!50] (13,4) rectangle (18,14);
+      \draw (13,4) -- (13,14);
+      \draw (18,4) -- (18,14);
+      \draw (15.5,15) node {code memory};
+    \end{scope}
+
+    \begin{scope}
+      \fill[gray!50] (13,0) rectangle (18,2);
+      \draw (13,0) -- (13,2);
+      \draw (18,0) -- (18,2);
+      \draw (15.5,3) node {heap memory};
     \end{scope}
@@
       \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,5) rectangle (18,6.5) node[pos=.5] {\footnotesize caller code};
     \end{onlyenv}
@@
       \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,5) rectangle (18,6.5) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,7) rectangle (18,8.5) node[pos=.5] {\footnotesize callee code};
       \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
@@
     \begin{onlyenv}<+-.(2)>
-      \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
+      \draw (13,1.5) edge[red,bend right,thick,->] (ret1);
     \end{onlyenv}
@@
     \begin{onlyenv}<+>
-      \draw[teal,very thick] (9,4.9) -- (12.5,5.9);
-      \draw[teal,very thick] (9,5.9) -- (12.5,4.9);
+      \draw[teal,very thick] (11.3,3) -- (13,2.8);
+      \draw[teal,very thick] (11.8,3.7) -- (12.4,2);
       \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
@@
     \begin{onlyenv}<+->
-      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      % \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
       \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
       \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
-      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=gray!30] (13,7) rectangle (18,8.5) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,9) rectangle (18,10.5) node[pos=.5] {\footnotesize callee code (1)};
     \end{onlyenv}
@@
     \begin{onlyenv}<+->
-      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
-      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      % \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      % \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
       \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
       \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
-      \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+      \draw[fill=gray!30] (13,9) rectangle (18,10.5) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,11) rectangle (18,12.5) node[pos=.5] {\footnotesize callee code (2)};
     \end{onlyenv}

### 2017-12-19 fully-abstract-comp/technical-report.tex
reordered a bit, so the helpful definitions and so on comes earlier.
Added FTLR
removed some rogue merge stuff

@@
 
+\subsubsection{Helpful functions, sets, and conventions}
+
+\[
+  \pi_\baddr(\vsc) =
+  \begin{cases}
+    \baddr & \text{if $\vsc = ((\_,\_),\baddr,\_,\_)$} \\ 
+    \pi_\baddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \stkptr{\_,\baddr,\_,\_}$}} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \retptrd(\_,(\baddr,\_))$}} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\baddr,\_,\_))$}}
+  \end{cases}
+\]
+\[
+  \pi_\eaddr(\vsc) =
+  \begin{cases}
+    \eaddr & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
+    \pi_\eaddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrd(\_,(\_,\eaddr))$}} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\_,\eaddr,\_))$}}
+  \end{cases}
+\]
+
+\[
+  \pi_\lin(\vsc) = 
+  \begin{cases}
+    \lin & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
+    \pi_\lin(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
+    \src{\linear} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
+    \src{\linear} & \src{\text{if $\vsc = \retptrd(\_,(\_,\_))$}} \\ 
+    \src{\lin} & \src{\text{if $\vsc = \retptrc(\_,(\_,\lin,\_,\_,\_))$}}
+  \end{cases}
+\]
+
+\dominique{4-12-2017: what about $\pi_a$?}
+\lau{12-12-2017: I have added them as I needed them in the LR. I have not needed $\pi_a$ yet.}
+\[
+  \updPcAddr{\Phi} \defeq 
+  \begin{cases}
+    \Phi\update{\pcreg}{((\perm,\lin),\baddr,\eaddr,\aaddr+1)} & \text{if $\Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$}\\
+    \failed & \totherwise
+  \end{cases}
+\]
+
+\[
+  \readAllowed{} \defeq \{ \rwx,\rw,\rx,\ro \}
+\]
+
+\[
+  \writeAllowed{} \defeq \{ \rwx, \rw \}
+\]
+
+\[
+  \isLinear{\cb} \defeq
+  \begin{cases}
+    \top & 
+    \arraycolsep=0pt
+    \begin{array}[t]{l}
+      \cb = ((\_,\linear),\_,\_,\_) \tor\\
+      \cb = \sealed{\_,\cb'} \tand \isLinear{\cb'} 
+    \end{array}\\
+    \src{\top} & 
+    \sourcecolor\left.
+    \arraycolsep=0pt
+    \begin{array}[t]{l}
+      \cb = \stkptr{\_,\_,\_,\_}\\
+      \cb = \retptrd(\_,\_)
+    \end{array}\right.\\
+    \bot & \totherwise
+  \end{cases}
+\]
+
+\[
+  \nonLinear{\cb} \defeq \neg \isLinear{\cb}
+\]
+
+\[
+  \linCons{w} \defeq
+  \begin{cases}
+    0 & \isLinear{w} \\
+    w & \totherwise
+  \end{cases}
+\]
+
+\[
+  \exec{\cb} \defeq 
+      \cb = ((\perm, \_), \_, \_, \_) \tand \perm \in \{\rwx,\rx\} 
+\]
+
+\[
+  \nonExec{\cb} \defeq \neg \exec{\cb}
+\]
+
+\[
+  \withinBounds{\vsc} \defeq 
+  \begin{cases}
+    \baddr \leq \aaddr \leq \eaddr & 
+    \arraycolsep=0pt
+    \begin{array}[t]{l}
+      \vsc = ((\_,\_),\baddr,\eaddr,\aaddr) \src{\tor}\\
+      \src{\vsc = \stkptr{\_,\baddr,\eaddr,\aaddr}}
+    \end{array}\\
+    \sigma_\baddr \leq \sigma_\aaddr \leq \sigma_\eaddr & \vsc = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \\
+    \bot & \totherwise
+  \end{cases}
+\]
+
+\[
+  \nonZero{w} \defeq
+  \begin{cases}
+    \bot & w \in \ints \tand w = 0 \\
+    \top & \totherwise
+  \end{cases}
+\]
+
+For convenience, we introduce the following notation:
+\[
+  \begin{array}{rcl}
+    \Phi(r) & \defeq & \Phi.\reg(r)
+  \end{array}
+\]
+where $r\in \RegName$.
+
+For $\rn$ which can be a register or an integer, we take
+\[
+  \Phi(\rn) = n
+\]
+to mean
+\[
+  \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$ and in either case $n \in \ints$}
+\]
+
 \subsubsection{Step relations}
@@
     \text{// split the stack at its current address - rstk done.}\\
-    \tgeta{r_{t1}}{\rstk}\\
-    \tsplit{\rstk}{\rretd}{\rstk}{r_{t1}}\\
+    \tgeta{\rtmp{1}}{\rstk}\\
+    \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}\\
     \text{// load the seal for the return pointer through the pc capability.}\\
-    \tmove{r_{t1}}{\pcreg}\\
-    \tsetatob{r_{t1}}{r_{t1}}\\
-    \tload{r_{t1}}{r_{t1}}\\
-    \tcca{r_{t1}}{o}\text{ // $o$ is the offset this call uses.}\\
+    \tmove{\rtmp{1}}{\pcreg}\\
+    \tsetatob{\rtmp{1}}{\rtmp{1}}\\
+    \tload{\rtmp{1}}{\rtmp{1}}\\
+    \tcca{\rtmp{1}}{o}\text{ // $o$ is the offset this call uses.}\\
     \text{// seal the used stack frame as the data part of the return pointer pair.}\\
-    \tcseal{\rretd}{\rretd}{r_{t1}}\\
+    \tcseal{\rretd}{\rretd}{\rtmp{1}}\\
     \text{// obtain the code part of the return pointer pair and seal it too.}\\
@@
     \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
-    \tcseal{\rretc}{\rretc}{r_{t1}}\\
+    \tcseal{\rretc}{\rretc}{\rtmp{1}}\\
     \text{// now clear temporary register and jump to the adversary.}\\
-    \tmove{r_{t1}}{0}\\
+    \tmove{\rtmp{1}}{0}\\
     \txjmp{r_1}{r_2}\\
@@
     \text{// check that the stack pointer is the same we handed out.}\\
-    \tgetb{r_{t1}}{\rstk}\\
-    \tminus{r_{t1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
-    \tmove{r_{t2}}{\pcreg}\\
-    \tcca{r_{t2}}{5} \text{ //magic number is the offset to fail} \\
-    \tjnz{r_{t2}}{r_{t1}} \\
-    \tcca{r_{t2}}{1} \text{ //magic number is the offset to after fail} \\
-    \tjmp{r_{t2}} \\
+    \tgetb{\rtmp{1}}{\rstk}\\
+    \tminus{\rtmp{1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
+    \tmove{\rtmp{2}}{\pcreg}\\
+    \tcca{\rtmp{2}}{5} \text{ //magic number is the offset to fail} \\
+    \tjnz{\rtmp{2}}{\rtmp{1}} \\
+    \tcca{\rtmp{2}}{1} \text{ //magic number is the offset to after fail} \\
+    \tjmp{\rtmp{2}} \\
     \tfail \\
@@
     \text{// clear temporary registers used}\\
-    \tmove{r_{t1}}{0}\\
-    \tmove{r_{t2}}{0}\\
+    \tmove{\rtmp{1}}{0}\\
+    \tmove{\rtmp{2}}{0}\\
     \text{// continue program after invocation.}
@@
 \]
-
-\subsubsection{Helpful functions, sets, and conventions}
-
-\[
-  \pi_\baddr(\vsc) =
-  \begin{cases}
-    \baddr & \text{if $\vsc = ((\_,\_),\baddr,\_,\_)$} \\ 
-    \pi_\baddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \stkptr{\_,\baddr,\_,\_}$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrd(\_,(\baddr,\_))$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\baddr,\_,\_))$}}
-  \end{cases}
-\]
-\[
-  \pi_\eaddr(\vsc) =
-  \begin{cases}
-    \eaddr & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
-    \pi_\eaddr(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrd(\_,(\_,\eaddr))$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\_,\eaddr,\_))$}}
-  \end{cases}
-\]
-
-\[
-  \pi_\lin(\vsc) = 
-  \begin{cases}
-    \lin & \text{if $\vsc = ((\_,\_),\_,\eaddr,\_)$} \\ 
-    \pi_\lin(c) & \text{if $\vsc = \sealed{\_,c}$} \\ 
-    \src{\linear} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
-    \src{\linear} & \src{\text{if $\vsc = \retptrd(\_,(\_,\_))$}} \\ 
-    \src{\lin} & \src{\text{if $\vsc = \retptrc(\_,(\_,\lin,\_,\_,\_))$}}
-  \end{cases}
-\]
-
-\dominique{4-12-2017: what about $\pi_a$?}
-\lau{12-12-2017: I have added them as I needed them in the LR. I have not needed $\pi_a$ yet.}
-\[
-  \updPcAddr{\Phi} \defeq 
-  \begin{cases}
-    \Phi\update{\pcreg}{((\perm,\lin),\baddr,\eaddr,\aaddr+1)} & \text{if $\Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$}\\
-    \failed & \totherwise
-  \end{cases}
-\]
-
-\[
-  \readAllowed{} \defeq \{ \rwx,\rw,\rx,\ro \}
-\]
-
-\[
-  \writeAllowed{} \defeq \{ \rwx, \rw \}
-\]
-
-\[
-  \isLinear{\cb} \defeq
-  \begin{cases}
-    \top & 
-    \arraycolsep=0pt
-    \begin{array}[t]{l}
-      \cb = ((\_,\linear),\_,\_,\_) \tor\\
-      \cb = \sealed{\_,\cb'} \tand \isLinear{\cb'} 
-    \end{array}\\
-    \src{\top} & 
-    \sourcecolor\left.
-    \arraycolsep=0pt
-    \begin{array}[t]{l}
-      \cb = \stkptr{\_,\_,\_,\_}\\
-      \cb = \retptrd(\_,\_)
-    \end{array}\right.\\
-    \bot & \totherwise
-  \end{cases}
-\]
-
-\[
-  \nonLinear{\cb} \defeq \neg \isLinear{\cb}
-\]
-
-\[
-  \linCons{w} \defeq
-  \begin{cases}
-    0 & \isLinear{w} \\
-    w & \totherwise
-  \end{cases}
-\]
-
-\[
-  \exec{\cb} \defeq 
-      \cb = ((\perm, \_), \_, \_, \_) \tand \perm \in \{\rwx,\rx\} 
-\]
-
-\[
-  \nonExec{\cb} \defeq \neg \exec{\cb}
-\]
-
-\[
-  \withinBounds{\vsc} \defeq 
-  \begin{cases}
-    \baddr \leq \aaddr \leq \eaddr & 
-    \arraycolsep=0pt
-    \begin{array}{l}
-      \vsc = ((\_,\_),\baddr,\eaddr,\aaddr) \src{\tor}\\
-      \src{\vsc = \stkptr{\_,\baddr,\eaddr,\aaddr}}
-    \end{array}\\
-    \sigma_\baddr \leq \sigma_\aaddr \leq \sigma_\eaddr & \vsc = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \\
-    \bot & \totherwise
-  \end{cases}
-\]
-
-\[
-  \nonZero{w} \defeq
-  \begin{cases}
-    \bot & w \in \ints \tand w = 0 \\
-    \top & \totherwise
-  \end{cases}
-\]
-
-For convenience, we introduce the following notation:
-\[
-  \begin{array}{rcl}
-    \Phi(r) & \defeq & \Phi.\reg(r)
-  \end{array}
-\]
-where $r\in \RegName$.
-
-For $\rn$ which can be a register or an integer, we take
-\[
-  \Phi(\rn) = n
-\]
-to mean
-\[
-  \text{either $n = \rn$ or $n = \Phi.\reg(\rn)$ and in either case $n \in \ints$}
-\]
-
-\subsubsection{Step Relation}
 
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
-    \right\} \cup \\
+    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    % \right\} \cup \\
 %
@@
 \]
+
+\subsection{Fundamental Theorem of Logical Relations}
+
+\begin{theorem}[FTLR]
+  \label{thm:ftlr}
+  \[
+    \begin{array}{l}
+      \forall n,W,W_D,W_C,\perm_C,\perm_D,\lin_C,\lin_D,\baddr_C,\baddr_D,\eaddr_C,\eaddr_D,\aaddr_C,\aaddr_D \ldotp \\
+      \quad \left(
+      \begin{array}{l}
+        W = W_D \oplus W_C \wedge \\
+        \perm_D \in \readAllowed{} \Rightarrow \npair{(\baddr_D,\eaddr_D)} \in \readCond{\lin_D,W_D} \wedge\\
+        \perm_D \in \writeAllowed{} \Rightarrow\npair{(\baddr_D,\eaddr_D)} \in \writeCond{\lin_D,W_D} \wedge\\
+        \perm_D \notin \{\rwx,\rx\} \wedge \\
+        \perm_C = \rx \wedge\\
+        \npair{(\{\rx\},\baddr_C,\eaddr_C)} \in \execCond{\lin_C,W_C}
+      \end{array}
+      \right)\\
+      \quad\Rightarrow 
+      \npair{\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
+                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray} \in \lre(W)
+
+    \end{array}
+  \]
+\end{theorem}
 

### 2017-12-19 presentations/illustrations/linear-prevention.tex
Clarify attack 3 prevention in linear slides a bit.

@@
   \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+  % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
 
@@
     \draw (18,0) -- (18,13);
-    \draw (15.5,14) node {heap memory};
+    \draw (15.5,14) node {code memory};
   \end{scope}
@@
 
-  \begin{onlyenv}<+->
+  \begin{onlyenv}<+-+(3)>
     \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
@@
 \note{Before the callee jumps, it is assumed that it has checked the global "max height" of its stack pointer!}
-  \begin{onlyenv}<+->
+  \begin{onlyenv}<+-+(4)>
     \capbrace[ret2]{(6.1,4.5)}{(6.1,7.5)}
@@
 
-  \begin{onlyenv}<+->
+  \begin{onlyenv}<+-+(1)>
     \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
@@
 \note{Callee (1) calls callee (2)}
-  \begin{onlyenv}<+->
+  \begin{onlyenv}<+-+(2)>
     \capbrace[ret3]{(6.1,7.5)}{(6.1,10.5)}
-    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-    \draw node[right=0cm of ret3] { ret-ptr (2) };
+    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr-d (2)} };
+    \draw node[right=0cm of ret3] { ret-ptr-d (2) };
   \end{onlyenv}
@@
     \inactsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
+  \end{onlyenv}
 
+  \begin{onlyenv}<.->
     \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
@@
   \begin{onlyenv}<+->
-    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=gray!30] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
     \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
@@
   \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
+    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
+    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (10,6) { Splice fails!};  
+  \end{onlyenv}
+
 \note{Splicing: fails because ends of capabilities don't meet. callee stack frame (1) cannot be empty (it will at least contain 42).}
-  \begin{onlyenv}<+->
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (8,7.5) {Caller tries to splice sp \\and ret-ptr-d, but fails!};  
-  \end{onlyenv}
+  % \begin{onlyenv}<+->
+  %   \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (8,7.5) {Caller tries to splice sp \\and ret-ptr-d, but fails!};  
+  % \end{onlyenv}
 \note{Before the caller makes another call/return, it will check that the stack height matches the global max stack height.}

### 2017-12-19 presentations/prisc18-01-2017-linear-caps/presentation.tex
Presentation for PriSC.

\documentclass{beamer}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}

\newcommand{\isdef}{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}

\title{Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
\institute{\inst{1}Aarhus University \qquad \inst{2}KU Leuven}
\date{PriSC 2018, Los~Angeles}

\input{../illustrations/preamble}
\begin{document}
\frame{\titlepage}

\begin{frame}
  \frametitle{A realistic secure compiler?}

  \vspace{-1cm}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % \draw<1-6> (-5,0) node {};
    % \draw (5,0) node {};
    % \draw (0,-14) node {};
    \draw (0,0) node[draw=black] (ir) {...};
    \draw node[draw=black,rounded corners,above=of ir] (source) {Source Language};
    \draw (source) edge[bend left, ->] (ir);
    \begin{onlyenv}<1-3>
      \draw node[draw=black,rounded corners,below=of ir] (target) {Target Language};
      \draw (ir) edge[bend left, ->] (target) ;
    \end{onlyenv}
    \begin{onlyenv}<2>
      \draw node[right=1cm of ir, line width=1mm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{onlyenv}
    \draw<3> node[right=1cm of target, line width=1mm, font=\fontfamily{ppl}\fontsize{1.5cm}{2cm}\selectfont] {?};
    \begin{visibleenv}<4-5>
      \draw (200:5) node[draw=black,rounded corners, text width=3.5cm] (commhar) {Commodity Hardware\\(Virtual Memory)};
      \draw (ir) edge[bend right=10, ->] (commhar) ;
      \draw (230:5) node[draw=black,rounded corners, text width=3.5cm] (wasm) {WebAssembly};
      \draw (ir) edge[bend left, ->] (wasm) ;
    \end{visibleenv}
    \begin{visibleenv}<4-6>
      \draw (280:5.8) node[draw=black,rounded corners, text width=3.5cm] (dtal) {Dependently-typed\\Assembly Language};
      \draw (ir) edge[bend left=20, ->] (dtal) ;
      \draw (335:8) node[draw=black,rounded corners, text width=3.5cm] (mph) {Micro-Policies Hardware};
      \draw (ir) edge[bend right=20, ->] (mph) ;
      \draw (350:7) node[draw=black,rounded corners, text width=3.5cm] (capmach) {Capability Machines};
      \draw (ir) edge[bend left=10, ->] (capmach) ;
    \end{visibleenv}
    \begin{visibleenv}<5>
      \draw node[below left=0cm of commhar, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
      \draw node[below left=0cm of wasm, font=\fontfamily{ppl}\fontsize{1cm}{1.5cm}\selectfont] {?};
    \end{visibleenv}
    \begin{visibleenv}<7>
      \draw node[draw=black,rounded corners,below=of ir] (capmach2) {Capability machines};
      \draw (ir) edge[bend left, ->] (capmach2) ;
    \end{visibleenv}
    \begin{visibleenv}<8->
      \draw node[draw=black,below=of ir] (ir2) {...};
      \draw (ir) edge[bend left, ->] (ir2) ;
      \draw node[draw=black,rounded corners,below=of ir2] (capmach3) {Capability machines};
      \draw (ir2) edge[bend left, ->] (capmach3) ;
    \end{visibleenv}
    \begin{visibleenv}<9>
      \draw node[rectangle,fit=(ir2) (capmach3),draw=red] (laststeps) {};
      \draw node[above right=0cm of laststeps,cloud,red] {Let's take a look!};
    \end{visibleenv}
  \end{tikzpicture}
  \only<5>{Fine-grained encapsulation?}
  \only<8>{Well-bracketed control flow, stack encapsulation?}
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

% \AtBeginSection[]
% {
%  \begin{frame}<beamer>
%  \frametitle{Outline}
%  \tableofcontents[currentsection]
%  \end{frame}
% }
\section{Stack and return pointers}
\input{../illustrations/traditional}

\begin{frame}<1>[label=sarps]
  \frametitle{Stack and return pointer security}
  \begin{itemize}
  \item Desirable properties:
    \begin{itemize}
    \item Well-bracketed control flow
    \item Stack frame encapsulation
    \end{itemize}
  \item How to enforce?
    \begin{itemize}
    \item Capabilities?
      \begin{itemize}
      \item<2-> Not enough: two attacks!
      \end{itemize}
    \item<3-> CHERI Local capabilities?
      \begin{itemize}
      \item Cannot leave registers
      \item<4-> Works, but... [Skorstengaard et al., ESOP 2018]
      \end{itemize}
    \item<5-> CHERI Linear capabilities?
      \begin{itemize}
      \item Non-copyable capabilities
      \item<6> Works perfectly?
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Stack and return capabilities}
\input{../illustrations/capabilities}

\section{Capabilities are not enough}
\againframe<2>{sarps}
\input{../illustrations/attacks}

\section{Local stack and return capabilities}
\againframe<3>{sarps}
\input{../illustrations/local-prevention}

\againframe<4>{sarps}
\section{Linear stack and return capabilities}
\againframe<5>{sarps}
\input{../illustrations/linear-prevention}
\againframe<6>{sarps}

\begin{frame}
  \frametitle{Formally?}

  \begin{itemize}
  \item<2-> Formulate correctness?
    \begin{itemize}
    % \item<3-> (CFI is too weak (right flows \emph{at the right time}))
    \item<3-> Idea: well-bracketed version of target assembly
    \item<3-> Compiler replaces call/returns with calling convention
    \item<3-> Fully abstract compilation
    \end{itemize}
  \item Prove correctness?
    \begin{itemize}
    \item<4-> Back-translation as embedding
    \item<4-> Cross-language step-indexed Kripke logical relation
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
  \item Enforcing well-bracketedness and stack frame encapsulation on a capability machine
  \item Local capabilities work [Skorstengaard et al, ESOP 2018]
  \item Linear capabilities work better!
    \begin{itemize}
    \item but don't exist... yet?
    \end{itemize}
  \item Correctness as fully abstract compilation from well-bracketed version of target language.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Thanks}
  \begin{center}
    \Large Questions?
  \end{center}
\end{frame}

\appendix
\begin{frame}
  \frametitle{Backup slide: The Awkward Example}
  \begin{align*}
    \tau &\isdef (unit \rightarrow unit) \rightarrow int\\
    e_1 &\isdef
          \left\{\begin{aligned}
            &\mathrm{let}~ x~ =~ \mathrm{ref}~ 0~ \mathrm{in}~\\
          &\lambda f\ldotp x \mathrel{:=} 0;~ f~\mathrm{unit}; x \mathrel{:=} 1; f~ \mathrm{unit}; !x
          \end{aligned}\right.\\
    e_2 &\isdef \lambda f\ldotp (f~\mathrm{unit}; f~\mathrm{unit}; 1)
  \end{align*}
\end{frame}

\end{document}

### 2017-12-19 fully-abstract-comp/technical-report.tex
.
Added conditions for the encodeType function used for the getType instruction.

@@
 These two functions need to be related in the following way:
-\lau{14-09-2017: Insert the relation.
+% Lau: 19-12-2017: Resolved the belove by adding the conditions.
+%\lau{14-09-2017: Insert the relation.
 My impression is that the the words specific to the source machine should have the same "type" as their implementation at the source level (I am not sure this is the case - maybe it is okay to just relate the values returned by the special words to the ones returned when used with the implementation).
 }
-\dominique{20-9-2017: I also think gettype for source-specific words should produce the same result than gettype for their implementation.}
+%\dominique{20-9-2017: I also think gettype for source-specific words should produce the same result than gettype for their implementation.}
+\begin{itemize}
+\item $\encType{}(((\_,\_),\_,\_,\_))$, $\encType{}(\seal{\_,\_,\_})$, $\encType{}(\sealed{\_,\_})$, and $\encType{}(i)$ where $i\in\ints$ are all destinct.
+\item For all $w \in \SealableCaps$ (only the words on the target machine), $\encType{}_{\var{trg}}(w) = \encType{}_{\var{src}}(w)$.
+\item Finally, 
+\[
+\encType{}_{\var{src}}(\src{\stkptr{\_,\_,\_,\_}}) = \encType{}_{\var{trg}}(((\_,\_),\_,\_,\_))
+\]
+ and 
+\[
+\encType{}_{\var{src}}(\src{\retptrd(\_,(\_,\_))}) = \encType{}_{\var{src}}(\src{\retptrc(\_,(\_,\_,\_,\_,\_))}) = \encType{}_{\var{trg}}(\sealed{\_,\_})
+\]
+\end{itemize}
+In English this means that each type of word is represented by a distinct value and that the tokens on the source machine has the type of the capability they represent on the target machine.
 \begin{align*}
@@
       \quad\Rightarrow 
-      \npair{\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
+      \npair{\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray} \in \lre(W)

### 2017-12-19 fully-abstract-comp/technical-report.tex
Added conditions for decoding and encoding function for permissions as well as encoding function for linearity

@@
 \newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
+\newcommand{\encLin}[1]{\plainfun{encoceLin}{#1}}
 \newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
@@
 \noindent\textbf{geta, getb, gete, getp, and getl}\\
-We assume function to encode and decode permissions as well as a function to encode and decode linearity.
+We assume functions to encode and decode permissions as well as a function to encode linearity.
 The functions are used implicitly when a permission or linearity is used in a place where they need to be a word.
-\lau{14-09-2017: Write more about this (types and assumptions)}
+%\lau{14-09-2017: Write more about this (types and assumptions)}
+
+Specifically for the permission function, $\encPerm{} : \Perm \fun \ints$ and $\decPerm{} : \ints \fun \Perm $ encodes and decodes permissions, respectively.
+Where $\decPerm{}$ is the left inverse of $\encPerm{}$, $\encPerm{}$ is injective and for all $\perm \in \Perm$ $\encPerm{\perm} \neq -1$ (as this is used as an error value).
+$\decPerm{}$ is surjective.
+
+For the linearity encoding, we make similar assumptions: $\encLin{} : \Linear \fun \ints$ encodes linearity.
+$\encLin{}$ is injective and for all $\lin \in \Linear$ $\encPerm{\lin} \neq -1$ (as this is used as an error value).
+As a capabilities linearity stays the same, we do not need a decoding function for linearity.
 \begin{align*}

### 2017-12-19 fully-abstract-comp/technical-report.tex
updated getl to return linear/normal for all words (as this is observable anyway). It will also return normal for ints.
some more updates to op sem. Added a missing case in seta2b

@@
       \cb = \stkptr{\_,\_,\_,\_}\\
-      \cb = \retptrd(\_,\_)
+      \cb = \retptrd(\_,\_)\\
+      \cb = \retptrd(\_,\linear,\_,\_,\_)
     \end{array}\right.\\
@@
                                    \begin{cases}
-                                     \updPcAddr{\Phi\updReg{r_1}{\lin}} & 
-                                     \arraycolsep=0pt
-                                     \begin{array}[t]{l}
-                                       \Phi(r_2) = ((\_,\lin),\_,\_,\_) \\
-                                       \quad \src{\tor \Phi(r_2) = \stkptr{\_,\_,\_,\_} \tand l = \linear}\\
+                                     \updPcAddr{\Phi\updReg{r_1}{\linear}} & 
+                                     \arraycolsep=0pt
+                                     \begin{array}[t]{l}
+                                       \isLinear{\Phi(r_2)}
                                      \end{array} \\
-                                     \updPcAddr{\Phi\updReg{r_1}{\normal}} & \Phi(r_2) = \seal{\_,\_,\_}\\
-                                     \updPcAddr{\Phi\updReg{r_1}{-1}} & \totherwise
+                                     \updPcAddr{\Phi\updReg{r_1}{\normal}} & \nonLinear{\Phi(r_2)}
                                    \end{cases}
 \end{align*}
+\lau{19-12-2017: Resolved the comments below. Seals are all normal and we allow linearity of sealed capabilities to be retrieved as it is observable anyway.}
 \lau{15-09-2017: Should getl do something for seals?}
@@
 \dominique{Linearity of sealed caps is already observable anyway, so no point in hiding it.
-Permissions of the sealed capability should be hidden, I think, mostly out of conservativity (hide as much as possible!).}
+Permissions of the sealed capability should be hidden, I think, mostly out of conservativity (hide as much as possible!).}\\
 \dominique{4-12-2017: Couldn't the definitions above use $\pi_b$, $\pi_e$, $\pi_l$ and $\pi_a$ and $\pi_p$ (the latter two to be added) ?}
@@
 \end{align*}
-Notice that in the case where we are moving a linear capability and $r = \rn$ the order of the updates matter.
+(Notice that in the case where we are moving a linear capability and $r = \rn$ the order of the updates matter.)
 
@@
                                           \Phi&\updReg{r_2}{w_2}\\
-                                              &\update{\mem.a}{w}
+                                              &\update{\mem.a}{\Phi(r_2)}
                                         \end{array}
@@
                                         \withinBounds{\Phi(r_1)} \tand \\
-                                        w = \Phi(r_2) \tand \\
-                                        w_2 = \linCons{w}
+                                        w_2 = \linCons{\Phi(r_2)}
                                       \end{array}
@@
                                         \Phi&\updReg{r_2}{w_2}\\
-                                            &\update{\ms_\stk.a}{w}
+                                            &\update{\ms_\stk.a}{\Phi(r_2)}
                                       \end{array}
@@
                                         \withinBounds{\Phi(r_1)} \tand \\
-                                        \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
-                                        w = \Phi(r_2) \\
-                                        w_2 = \linCons{w}
+                                        \aaddr \in \dom(\Phi.\ms_\stk) \tand \\ % Lau: this condition seems odd as there should never be a capability for a piece of memory that is unreachable.
+                                        w_2 = \linCons{\Phi(r_2)}
                                       \end{array}\right.\\
@@
                                       \begin{array}{rl}
-                                        \Phi&\updReg{r_1}{w}\\
-                                            &\update{\mem.a}{w_2}
+                                        \Phi&\update{\mem.a}{w_2}\\
+                                            &\updReg{r_1}{w}
                                       \end{array}\right)& 
@@
                                       \begin{array}{rl}
-                                        \Phi&\updReg{r_1}{w}\\
-                                            & \update{\mem_\stk.a}{w_2}
+                                        \Phi& \update{\ms_\stk.a}{w_2}\\
+                                            &\updReg{r_1}{w}
                                       \end{array}\right)\right.
@@
                                       \withinBounds{\Phi(r_2)} \tand \\
-                                      \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
+                                      \aaddr \in \dom(\Phi.\ms_\stk) \tand \\ % Lau: this condition seems odd as there should never be a capability for a piece of memory that is unreachable.
                                       w = \Phi.\ms_\stk(a) \tand \\
@@
 \noindent\textbf{cca}\\
-\dominique{4-12-2017: The phrase $\text{either $n = \rn$ or $n = \Phi.\reg(\rn)$ and in either case $n \in \ints$ and}$ is repeated a lot in the semantics.
+% The following comment has been resolved by implementing the suggestion.
+% \dominique{4-12-2017: The phrase $\text{either $n = \rn$ or $n = \Phi.\reg(\rn)$ and in either case $n \in \ints$ and}$ is repeated a lot in the semantics.
 Perhaps define an abbreviation as $\Phi(\rn) = n$?}
-
 \emph{Change Current Address}
-\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
+\\\lau{15-09-2017: This is the old lea. I changed the name, so it actually reflects what the instruction does. We might also want to add a real lea at some point.}
 \begin{align*}
@@
                                     \\
-                                    \updPcAddr{\Phi\updReg{r}{c}} &  
+                                    \updPcAddr{\Phi\updReg{r}{s}} &  
                                     \arraycolsep=0pt
@@
                                       \Phi(r) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \tand \\
-                                      c = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma + n}
+                                      s = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma + n}
                                     \end{array}
@@
 \begin{align*}
-  \sem{\trestrict{r_1}{r_2}{\rn}} = &
+  \sem{\trestrict{r_1}{r_2}{\rn}}(\Phi) = &
                                       \begin{cases}
@@
                                     \end{array} \\
+\sourcecolor\left.
+                                  \updPcAddr{}\left(
+                                    \arraycolsep=0pt
+                                    \begin{array}{rl}
+                                    \Phi&\updReg{r_2}{0}\\
+                                        &\updReg{r_1}{c}
+                                    \end{array} \right)
+\right.
+&
+\sourcecolor\left.
+                                    \arraycolsep=0pt
+                                    \begin{array}[t]{l}
+                                      \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\_} \tand \\
+                                      c = \stkptr{\perm,\baddr,\eaddr,\baddr}
+                                    \end{array}\right. \\
                                     \failed & \totherwise
@@
 \begin{align*}
-  \sem{\tsplit{r_1}{r_2}{r_3}{\rn_4}} = &
+  \sem{\tsplit{r_1}{r_2}{r_3}{\rn_4}}(\Phi) = &
                                \begin{cases}
@@
 \begin{align*}
-  \sem{\tsplice{r_1}{r_2}{r_3}} = &
+  \sem{\tsplice{r_1}{r_2}{r_3}}(\Phi) = &
                               \begin{cases}

### 2017-12-20 fully-abstract-comp/technical-report.tex
updated K-rel
updated K-rel
mend
Some more minor updates to the op sem. Please read the new question under interpretation of call.

@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_1) = \sealed{\sigma,c_\code} \tand \\
-                                       \Phi(r_2) = \sealed{\sigma,c_\data} \tand \\
+                                       \Phi(r_1) = \sealed{\sigma_1,c_\code} \tand \\
+                                       \Phi(r_2) = \sealed{\sigma_2,c_\data} \tand \\
+                                       \sigma_1 = \sigma_2 \tand \\
                                        \nonExec{c_\data} \tand \\
@@
                                        \begin{array}[t]{l}
-                                         \Phi(r_1) = \retptrc(\sigma,\_) \tand \\
-                                         \Phi(r_2) = \retptrd(\sigma,(\aaddr_\stk,\_)) \tand \\
+                                         \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
+                                         \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\_)) \tand \\
+                                         \sigma_1 = \sigma_2 \tand\\
                                          \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
@@
 \lau{25-10-2017: I have forgotten why we save $e_\stk$ on the stack and whether this is still necessary (it at least prevent the empty case, but I do not recall if it is necessary for other reasons).}
-This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack.
-We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}
+\lau{20-12-2017: At some point, I added an offset annotation $o$ to all calls.
+I believe this removes the need for a function that annotate seals.
+It will still have to be checked that all calls in a program uses different offsets, but I suspect this can be done as some assumption in what we want to prove.}
+\lau{20-12-2017: Until this day, we had a $\sealAss{}$ function: This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack.
+We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}}
 \begin{align*}
@@
                                                         &\updReg{\rretc}{\retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))}\\
-                                                        &\updReg{\rretd}{\retptrd(\sigma_?,(\aaddr_\stk,\eaddr_\stk))}
+                                                        &\updReg{\rretd}{\retptrd(\sigma_?,(\aaddr_\stk,\eaddr_\stk))}\\
+                                                        &\updReg{\rtmp{1}}{0}
                                                  \end{array}
@@
                                                    \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk - 1]} \tand \\
-                                                   c_\stk = \stkptr{\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
+                                                   c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
                                                    \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
@@
                                                    \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
-                                                   \mem(\baddr) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}\tand\\
+                                                   \mem(\baddr) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
+                                                   \sigma_? = \sigma_\aaddr + o \tand\\
                                                    \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
-                                                   \sigma_? = \sealAss{a} \tand\\
                                                    w_1 = \linCons{\Phi(r_1)} \tand \\
                                                    w_2 = \linCons{\Phi(r_2)}
-                                                 \end{array}
+                                                 \end{array}\\
+                                                 \failed & \totherwise
                                                \end{cases}
@@
 \lau{12-12-2017: Correct. Please feel free to correct these things.}
+\lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$.
+There is, however, no check in the implementation of call that ensures this.
+That says, we are going to have an assumption that the stack pointer base address has been checked before all calls (i.e., that it is $\stkb$) which would indeed mean that it is a stack pointer (assuming that the address $\stkb$ is a stack address to begin with).
+Why did we not want to check the stack end address here?
+And if we do not check it as part of the call implementation, do we then want to match the call implementation better?}
 
@@
       \qquad\memSat[n']{\stpair[.]{\ms_S,(\opc,\ms_{\stk,f}) :: \stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
-      \qquad\opc = c_c \wedge\\
+      \qquad\opc = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c+12) \wedge\\
       \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\

### 2017-12-20 fully-abstract-comp/technical-report.tex
more small changes to the section about incstruction interpretation.

@@
 \end{figure}
-\lau{13-09-2017: Note that most of the discussion belove is now captured in Section~\ref{sec:linear-cap}}
+\lau{13-09-2017: Note that most of the discussion below is now captured in Section~\ref{sec:linear-cap}}
 \lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear.
@@
 \begin{itemize}
-\item $\encType{}(((\_,\_),\_,\_,\_))$, $\encType{}(\seal{\_,\_,\_})$, $\encType{}(\sealed{\_,\_})$, and $\encType{}(i)$ where $i\in\ints$ are all destinct.
+\item $\encType{}(((\_,\_),\_,\_,\_))$, $\encType{}(\seal{\_,\_,\_})$, $\encType{}(\sealed{\_,\_})$, and $\encType{}(i)$ where $i\in\ints$ are all distinct.
 \item For all $w \in \SealableCaps$ (only the words on the target machine), $\encType{}_{\var{trg}}(w) = \encType{}_{\var{src}}(w)$.
@@
 \dominique{4-12-2017: Couldn't the definitions above use $\pi_b$, $\pi_e$, $\pi_l$ and $\pi_a$ and $\pi_p$ (the latter two to be added) ?}
-\lau{12-12-2017: $\pi_\_$ also projects from sealed capabilities, so the case of sealed capabilities would have to be explicitely excluded (with getl as the exception).}
+\lau{12-12-2017: $\pi_\_$ also projects from sealed capabilities, so the case of sealed capabilities would have to be explicitly excluded (with getl as the exception).}
 
@@
                                                & \updReg{\rdata}{0}\\
-                                               & \updReg{\rstk}{c_\stk}
+                                               & \updReg{\rstk}{c_\stk}\\
+                                               & \updReg{\rtmp{1}}{0}\\
+                                               & \updReg{\rtmp{2}}{0}
                                        \end{array}\right.&
@@
                                     \begin{array}{rl}
-                                      \Phi&\updReg{r_1}{\vsc}\\
-                                          &\updReg{r_2}{w_2}
+                                      \Phi&\updReg{r_2}{w_2}\\
+                                          &\updReg{r_1}{\vsc}
                                     \end{array}\right)
@@
 In other words, splice and split should not break linearity.
+%Alternative way to phrase the same thing: take the set of addresses governed by a linear capability.
+After a split, the union of the two sets that consist of the addresses the two capabilities govern should be equal to the set of addresses the original capability governed.
+Further, the two sets of addresses two capabilities uptained from a split govern should be disjoint.
 \end{enumerate}
@@
                                  \begin{array}{rl}
-                                   \Phi&\updReg{r_3}{w}\\
-                                       &\updReg{r_1}{c_1}\\
+                                   \Phi&\updReg{r_1}{c_1}\\
                                        &\updReg{r_2}{c_2}
@@
                                    c_1 = \seal{\sigma_\baddr,n,\sigma} \tand \\
-                                   c_2 = \seal{n+1,\sigma_\eaddr,\sigma} \tand \\
-                                   w = \linCons{\Phi(r_3)}
+                                   c_2 = \seal{n+1,\sigma_\eaddr,\sigma}         
                                  \end{array}\\
@@
                                 \begin{array}{r l}
-                                  \Phi&\updReg{r_2}{w_2}\\
-                                      &\updReg{r_3}{w_3}\\
-                                      &\updReg{r_1}{c}
+                                  \Phi&\updReg{r_1}{c}
                                 \end{array}\right)
@@
                                   \sigma_{\baddr,3} \leq \sigma_{\eaddr,3} \tand \\
-                                  c = \seal{\sigma_{\baddr,2},\sigma_{\eaddr,3},\sigma} \tand \\
-                                  w_2 = \linCons{\Phi(r_2)} \tand \\
-                                  w_3 = \linCons{\Phi(r_3)} \\
+                                  c = \seal{\sigma_{\baddr,2},\sigma_{\eaddr,3},\sigma}
                                 \end{array}\\
@@
                                                    \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
-                                                   \stk' = (\opc,\ms_{\stk,\priv},\aaddr_\stk) :: \stk \tand\\
+                                                   \stk' = (\opc,\ms_{\stk,\priv}) :: \stk \tand\\
                                                    \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
@@
 
-Eventhough we call them linear, they are more like \emph{affine} capabilities because our current operational semantics allows for them to be overwritten.
+Even though we call them linear, they are more like \emph{affine} capabilities because our current operational semantics allows for them to be overwritten.
 We choose to stick with the name as the term linear is also being misused by others for instance in Rust.

### 2018-01-07 fully-abstract-comp/technical-report.tex
Typo in execCond and correction to P-relation: pair of data now trivially okay. World needs to be split up
Added definition for terminating computation
Remove obsolete comment.

@@
 \end{align*}
+
+\paragraph{Termination computations}
+We use the following notation to write that a configuration $\Phi$ successfully terminates:
+\[
+  \Phi\term[i] \defeq \Phi \nstep[i] \halted
+\]
+if we just know $\Phi$ terminates in some number of steps, then we write
+\[
+  \Phi\term \defeq \exists i \ldotp \term[i]
+\]
 
@@
 \end{align*}
-\dominique{4-12-2017: $e_\stk$ is no longer stored in $\ms_{\stk,\priv}$ in updated call implementation.}
-\lau{12-12-2017: Correct. Please feel free to correct these things.}
 \lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$.
@@
 \[
-  \lrp(W) = \left\{
-    \npair{ \overline{\stpair[.]{c}{c'}} } \middle| 
+  \lrp(W) =
+    \left\{
+    \npair{ [\stpair[.]{c_1}{c'_1}, \dots, \stpair[.]{c_k}{c'_k}] } \middle| 
     \begin{array}{l}
-      \text{for $\stpair[.]{c}{c'} \in \overline{\stpair[.]{c}{c'}}$} \\
-      \src{\baddr} = \pi_\baddr(\src{c}) \wedge \src{\eaddr} = \pi_\eaddr(\src{c}) \wedge 
-      \baddr = \pi_\baddr(c) \wedge \eaddr = \pi_\eaddr(c) \wedge \\
-      \src{\lin} = \pi_\lin(\src{c}) = \pi_\lin(c) \\
-      \quad \exists r \in \addressable{l,W} \ldotp\\
-      \qquad \forall \npair{\stpair{\ms}{\ms}} \in W(r).H \; W(r).s \; \xi^{-1}(W)  \ldotp \\
-      \qquad \quad \dom(\src{\ms_S}) \subseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \subseteq [\baddr,\eaddr]      
+      \exists S : \{1,\dots,k\} \fun \World \ldotp \\
+      \quad W = \bigoplus_{i\in\{1,\dots,k\}} S(i)\wedge \\
+      \quad\text{for $i \in \{1,\dots,k\}$} \\
+      \qquad\stpair[.]{c_i}{c'_i} \in \ints \times \ints \vee \\
+      \qquad\src{\baddr} = \pi_\baddr(\src{c_i}) \wedge \src{\eaddr} = \pi_\eaddr(\src{c_i}) \wedge \\
+      \qquad\baddr = \pi_\baddr(c_i') \wedge \eaddr = \pi_\eaddr(c_i') \wedge \\
+      \qquad\src{\lin} = \pi_\lin(\src{c_i}) = \pi_\lin(c_i') \wedge \\
+      \qquad \exists r \in \addressable{l,S(i)} \ldotp\\
+      \qquad \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; S(i)(r).s \; \xi^{-1}(S(i))  \ldotp \\
+      \qquad \qquad \dom(\src{\ms_S}) \subseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \subseteq [\baddr,\eaddr]      
     \end{array}
@@
       \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
-      \qquad \npair[n']{[\stpair[.]{((\perm,lin),\baddr,\eaddr,\aaddr)}{((\perm,lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}]} \in \lrp(W' \oplus W'') \Rightarrow \\
-      \quad \qquad \npair[n']{\stpair[.]{((\perm,lin),\baddr,\eaddr,\aaddr)}{((\perm,lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
+      \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}]} \in \lrp(W' \oplus W'') \Rightarrow \\
+      \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
     \end{array}
@@
         \perm_C = \rx \wedge\\
-        \npair{(\{\rx\},\baddr_C,\eaddr_C)} \in \execCond{\lin_C,W_C}
+        \npair{(\baddr_C,\eaddr_C)} \in \readCond{\lin_C,W_C}
       \end{array}

### 2018-01-12 fully-abstract-comp/technical-report.tex
Wrote up suggestion for updated LR. Removed K-relation and wrote new cases for return capabilities (Future world relation still to be updated.)

@@
 We partly do not allow this out of convenience as it makes the implementation of call simpler.
-We do not need empty capabilities as they have no semantic value in the sense that they allow you to do no more than a piece of data.
+We do not need empty capabilities as they have no semantic value in the sense that they allow you to do essentially the same as a piece of data.
 \begin{align*}
@@
 \end{align*}
+\dominique{4-12-2017: $e_\stk$ is no longer stored in $\ms_{\stk,\priv}$ in updated call implementation.}
+\lau{12-12-2017: Correct. Please feel free to correct these things.}
 \lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$.
@@
       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair[.]{k_{c,S},k_{d,S}}{k_{c,T},k_{d,T}}} \in \lrk(W_\lrk) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rretd,\rretc,\rdata\})(W_\lrr) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_\lrr) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
 \right\}
-\]
-
-\[
-  \begin{array}{rl}
-  \lrk(W) = &\left\{ \npair{\left(\array{l}\src{\retptrc(\sigma,(\perm,\lin,\baddr_c,\eaddr_c,\aaddr_c))},\\\src{\retptrd(\sigma,(\aaddr_d,\eaddr_d))},\\k_{c,T},k_{d,T}\endarray\right)} \middle|
-    \begin{array}{l}
-      k_{d,T} = \sealed{\sigma,c_d} \wedge\\
-      k_{c,T} = \sealed{\sigma,c_c} \wedge\\
-      c_c = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c) \wedge\\
-      c_d = ((\rw,\linear),\aaddr_d,\eaddr_d,\aaddr_d-1) \wedge\\
-      \npair[n]{[\stpair[.]{\retptrc(\sigma,\_)}{k_{c,T}},\stpair[.]{\retptrd(\sigma,\_)}{k_{d,T}}]} \in \lrp(W) \wedge \\
-      \forall n' \leq n, W' \pubft W \ldotp \\
-      \quad \forall W_\lrr, W_\lrm \ldotp  \\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,(\opc,\ms_{\stk,f}) :: \stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
-      \qquad\opc = ((\perm,\lin),\baddr_c,\eaddr_c,\aaddr_c+12) \wedge\\
-      \qquad\dom(\ms_{\stk,f}) = [\aaddr_d,\eaddr_d] \wedge\\
-      \qquad\npair[n']{\left(\array{l}\src{\reg_S\update{\rstk}{\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}}},\\
-                               {\reg_T\update{\rstk}{((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)}}\endarray \right)}\in \lrr(\{r_\stk\})(W_\lrr) \wedge\\ 
-      \qquad\npair[n']{[(\opc,\opc)]} \in \lrp(W')\\
-      \qquad\Rightarrow\\
-      \qquad \npair[n']{\left(\arraycolsep=0pt\array{l}
-                          \left(\arraycolsep=0pt\array{l}
-                            \src{\ms_S,\reg_S\update{\pcreg,\rstk}{\opc,\stkptr{\rw,\stkb,\eaddr_d,\aaddr_d}},}\\
-                            \src{\stk, \ms_\stk \uplus \ms_{\stk,f}}
-                          \endarray \right) \\
-                          \left(\ms_T,\reg_T\update{\pcreg,\rstk}{\opc,((\rw,\linear),\stkb,\eaddr_d,\aaddr_d)} \right)
-                        \endarray \right)} \in \lro(W'\oplus W_\lrr \oplus W_\lrm)
-    \end{array}
-  \right\}
-% \\ \cup
-%     & \left\{ \npair{\left( k_{c,S},k_{d,S},k_{c,T},k_{d,T} \right)} \middle|
-%       \begin{array}{l}
-%         \pi_\sigma(k_{c,S}) = \pi_\sigma(k_{d,S}) \wedge \pi_\sigma(k_{c,T}) = \pi_\sigma(k_{d,T}) \wedge \npair{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \\
-%         \forall n' \leq n, W' \pubft W \ldotp \\
-%         \quad \npair[n']{(k_{c,S},k_{c,T}),(k_{d,S},k_{d,T})} \in \lrp(W) \Rightarrow\\
-%         \qquad \forall \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W'_\lrr) \ldotp 
-%                \forall \memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W'_\lrm} \ldotp\\
-%         \qquad \quad \npair[n']{\left(
-%         \arraycolsep=0pt\array{l}
-%           (\src{\ms_S},\src{\reg_S}\update{\pcreg,\rdata}{\pi_\vsc(\src{k_{c,S}}),\pi_\vsc(\src{k_{d,S}})},\src{\stk},\src{\ms_\stk}),\\
-%           (\ms_T,\reg_T\update{\pcreg,\rdata}{\pi_\vsc(k_{c,T}),\pi_\vsc(k_{d,T})})
-%         \endarray\right)} \in \lro(W' \oplus W'_\lrr \oplus W'_\lrm)
-%       \end{array}
-%      \right\}
-  \end{array}
 \]
@@
     \begin{array}{l}
-      \vsc_C = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge \\
-      \forall (\src{c},c') \in H_\sigma(\sigma) \ldotp \\
-      \quad \exists r \in \addressable{\linear,W} \ldotp \\
-      \qquad  \exists \stpair{\ms}{\ms} \ldotp \\
-      \qquad \quad \dom(\src{\ms_S}) = \dom(\ms_T) = [\pi_\baddr(c), \pi_\eaddr (c)] \wedge \stareg{\spao} \nequal W(r) \wedge \\
-      \qquad \qquad \forall W' \privft W \ldotp \\
-      \qquad \qquad \quad \exists r', W'(r') = W(r), \exists r'', W'(r'') = \stdreg{[\stkb,\eaddr]}{\spao} \ldotp \\
-      \qquad \qquad \qquad { \color{red} \npair{(\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)}, \src{c}, \sealed{\sigma,\vsc_c}, c')}} \in \lrk(W')
+      (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
+      \npair{(\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})} \in \lrp(W)
     \end{array}
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,\vsc_D}\endarray\right)} \;\middle| \;
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,(\baddr,\eaddr))},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)} \;\middle| \;
     \begin{array}{l}
-      \vsc_D = ((\rw,\linear),\baddr,\eaddr,\baddr-1) \wedge\\
-      \exists r \in \addressable{\linear,W} \ldotp \\
-      \quad  \exists \stpair{\ms}{\ms} \ldotp \\
-      \qquad \dom(\src{\ms_S}) = \dom(\ms_T) = [\baddr, \eaddr] \wedge \stareg{\spao} \nequal W(r) \wedge \\
-      \qquad \forall W' \privft W \ldotp \\
-      \qquad \quad \exists r', W'(r') = W(r), \exists r'', W'(r'') = \stdreg{[\stkb,\eaddr]}{\spao} \ldotp \\
-      \qquad \qquad { \color{red} \forall (\src{c}, c') \in H_\sigma (\sigma) \ldotp \npair{\src{c,\retptrd(\sigma,\baddr,\eaddr)}, c' , \sealed{\sigma,\vsc_D}} \in \lrk(W') }%\text{\lau{This needs some work - specifically what i c and c' and do we need to seal them?}}
-
-%    \forall n'<n, W' \pubft W, W_\lrv \npair[n']{\stpair[.]{\retptrc(\sigma,\_)}{\sealed{\sigma,\vsc_C}}} \in \lrv(W_\lrv) \ldotp\\
-%      \npair[n']{\src{\retptrc(\sigma,\_)},\src{\retptrd(\sigma,\_)},\sealed{\sigma,\vsc_C},\sealed{\sigma,\vsc_D}} \in \lrk(W' \oplus W_\lrv)
+      \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
+      \forall n' \leq n \ldotp \forall W' \privft W \ldotp \\
+      \quad \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
+      \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
+      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}{W_\lrm}} \wedge \\
+      \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata,\rstk,\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
+      \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
+      \quad \reg_T(r_\stk) = ((\rw,\linear),\stkb,\baddr-1,\_) \wedge \\
+      \quad W'(r_\var{free}) = \stdreg{[\stkb,\baddr-1]}{\spa} \wedge \\
+      \quad W'(r_\var{priv}) = \stareg[\src{\ms_f},\ms_f]{\spao} \wedge \\
+      \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
+      \quad W_\var{new} = (W' \uplus W_\lrr \uplus W_\lrm \uplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
+      \qquad \Rightarrow \\
+      \qquad \quad \npair[n']{\left(\array{l}(\ms_S,\reg_S\updReg{\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}}{\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0},\stk,\ms_\stk \uplus \ms_f),\\
+                                             (\ms_T,\reg_S\updReg{\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}}{((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?)})
+                                    \endarray\right)} \in \lro(W_\var{new})
     \end{array} 
@@
 \]
+where $H_\sigma$ is a function from seals to the set of pairs of sealable capabilities that can be sealed with $\sigma$.
 \lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}

### 2018-01-12 presentations/prisc18-01-2017-linear-caps/presentation.tex
Small fix in the PriSC presentation.

@@
 \date{PriSC 2018, Los~Angeles}
+
+\tikzset{
+    gray on/.style={alt=#1{gray}{}},
+    alt/.code args={<#1>#2#3}{%
+      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
+    },
+}
 
@@
     \end{visibleenv}
-    \begin{visibleenv}<4-6>
-      \draw (280:5.8) node[draw=black,rounded corners, text width=3.5cm] (dtal) {Dependently-typed\\Assembly Language};
-      \draw (ir) edge[bend left=20, ->] (dtal) ;
-      \draw (335:8) node[draw=black,rounded corners, text width=3.5cm] (mph) {Micro-Policies Hardware};
-      \draw (ir) edge[bend right=20, ->] (mph) ;
+    \begin{visibleenv}<4-7>
+      \draw (280:5.8) node[gray on=<7>,draw=black,rounded corners, text width=3.5cm] (dtal) {Dependently-typed\\Assembly Language};
+      \draw (ir) edge[gray on=<7>,bend left=20, ->] (dtal) ;
+      \draw (335:8) node[gray on=<7>,draw=black,rounded corners, text width=3.5cm] (mph) {Micro-Policies Hardware};
+      \draw (ir) edge[gray on=<7>,bend right=20, ->] (mph) ;
       \draw (350:7) node[draw=black,rounded corners, text width=3.5cm] (capmach) {Capability Machines};
@@
     \end{visibleenv}
-    \begin{visibleenv}<7>
+    \begin{visibleenv}<8>
       \draw node[draw=black,rounded corners,below=of ir] (capmach2) {Capability machines};
@@
     \end{visibleenv}
-    \begin{visibleenv}<8->
+    \begin{visibleenv}<9->
       \draw node[draw=black,below=of ir] (ir2) {...};
@@
     \end{visibleenv}
-    \begin{visibleenv}<9>
+    \begin{visibleenv}<10>
       \draw node[rectangle,fit=(ir2) (capmach3),draw=red] (laststeps) {};
@@
   \only<5>{Fine-grained encapsulation?}
-  \only<8>{Well-bracketed control flow, stack encapsulation?}
+  \only<9>{Well-bracketed control flow, stack encapsulation?}
 \end{frame}
@@
       \begin{itemize}
-      \item<2-> Not enough: two attacks!
+      \item<2-> Not enough: three attacks!
       \end{itemize}
@@
     % \item<3-> (CFI is too weak (right flows \emph{at the right time}))
-    \item<3-> Idea: well-bracketed version of target assembly
-    \item<3-> Compiler replaces call/returns with calling convention
+    \item<4-> Idea: well-bracketed version of target assembly
+    \item<4-> Compiler replaces call/returns with calling convention
     \item<3-> Fully abstract compilation
@@
     \begin{itemize}
-    \item<4-> Back-translation as embedding
-    \item<4-> Cross-language step-indexed Kripke logical relation
+    \item<5-> Back-translation as embedding
+    \item<5-> Cross-language step-indexed Kripke logical relation
     \end{itemize}

### 2018-01-17 esop-2018/paper.tex
anonymous version
A couple of changes prompted by ESOP reviewer 1

@@
   we provide a logical relation that semantically captures the guarantees
-  provided by the hardware (a form of capability safety). These results are not
-  tied to our calling convention and can be used to reason about arbitrary
-  programs.
+  provided by the hardware (a form of capability safety). The logical relation
+  is not specific to our calling convention and can be used to reason about
+  arbitrary programs.
 \end{abstract}
@@
 apply the well-studied techniques of step-indexed Kripke logical relations for
-reasoning about code on a representative capability machine code with local
+reasoning about code on a representative capability machine with local
 capabilities in general and correctness and security of the calling convention
@@
   \end{align*}
-  \begin{tabular}{|c|p{3.4cm}|p{7.3cm}|}
+  \begin{tabular}{|c|p{3.4cm}|>{\raggedright\arraybackslash}p{7.3cm}|}
     \hline

### 2018-01-17 esop-2018/paper.tex
more reviewer 1 fixes.

@@
 
-In order to have a reasonably realistic system, we use a simple model of
-linking, so programs can have access to other programs through a linking table.
-We also assume malloc to be part of the trusted computing base satisfying a
-certain specification. Malloc and linking tables are described further in the
-next section, but we refer to the technical appendix~\cite{technical_appendix}
-for full details.
+In order to have a reasonably realistic system, we use a simple model of linking
+where a program has access to a linking table that contains capabilities for
+other programs.  We also assume malloc to be part of the trusted computing base
+satisfying a certain specification. Malloc and linking tables are described
+further in the next section, but we refer to the technical
+appendix~\cite{technical_appendix} for full details.
 
@@
 
-\textbf{Reusable macro instructions} We define a number of reusable
-macros capturing the calling convention and other conveniences. All
-macros that use the stack assume a stack pointer in register
-$r_\stk$. The macro \texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the
-capability related to $\var{name}$ from the linking table and stores
-it in register $r$.  The macros \texttt{\footnotesize{push $r$}} and
-\texttt{\footnotesize{pop $r$}} work as one would
-expect. The macro \texttt{\footnotesize{prepstk $r$}} is used when a callback is
-invoked by the adversary and prepares the received stack pointer by
-checking that it has permission $\rwlx$.  The macro \texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}}
-jumps to the capability in register $r$ in the manner described
-above. That is, it pushes local state (the contents of registers
-$\overline{r_{\var{priv}}}$) and the activation record (return code,
-return pointer, stack pointer) to the stack, creates an $\entry$
-return pointer, restricts the stack pointer, clears the unused part of
-the stack, clears the necessary registers and jumps to $r$.  Upon
+\textbf{Reusable macro instructions} We define a number of reusable macros
+capturing the calling convention and other conveniences. All macros that use the
+stack assume a stack pointer in register $r_\stk$. The macro
+\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
+$\var{name}$ from the linking table and stores it in register $r$.  The macros
+\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} add and
+remove elements from the stack. The macro \texttt{\footnotesize{prepstk $r$}} is
+used when a callback is invoked by the adversary and prepares the received stack
+pointer by checking that it has permission $\rwlx$.  The macro
+\texttt{\footnotesize{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
+capability in register $r$ in the manner described above. That is, it pushes
+local state (the contents of registers $\overline{r_{\var{priv}}}$) and the
+activation record (return code, return pointer, stack pointer) to the stack,
+creates an $\entry$ return pointer, restricts the stack pointer, clears the
+unused part of the stack, clears the necessary registers and jumps to $r$.  Upon
 return, the private state is restored. The macro \texttt{\footnotesize{mclear
-    $r$}} clears all the memory the capability in register $r$ has
-authority over.  The macro \texttt{\footnotesize{rclear $\var{regSet}$}} clears
-all the registers in $\var{regSet}$. The macro \texttt{\footnotesize{reqglob
-    $r$}} checks whether the word in register $r$ is a global
-capability. The macro \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}}
-allocates a closure where $r$ points to the closure's code and a new
-environment is allocated (using malloc) where the contents of
-$\overline{r_i}$ is stored. In the code referred to by $r$, an
-implicit fetch happens when an instruction refers to $x_i$.
+    $r$}} clears all the memory the capability in register $r$ has authority
+over.  The macro \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the
+registers in $\var{regSet}$. The macro \texttt{\footnotesize{reqglob $r$}}
+checks whether the word in register $r$ is a global capability. The macro
+\texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}} allocates a closure
+where $r$ points to the closure's code and a new environment is allocated (using
+malloc) where the contents of $\overline{r_i}$ is stored. In the code referred
+to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
 
@@
 \emph{Modularity} It is important that our calling convention is modular, i.e.\
-we do not assume that our code is specially privileged w.r.t.\ the adversary, and
-they can apply the same measures to protect themselves from us as we do to
+we do not assume that our code is specially privileged w.r.t.\ the adversary,
+and they can apply the same measures to protect themselves from us as we do to
 protect ourselves from them. More concretely, the requirements we have on
@@
 store the old stack pointer, but the adversary's return pointers are also
-allowed to be local. Adversary callbacks on the other hand are required to
-be global but the callbacks we construct ourselves are allocated on the heap and
-also global.
+allowed to be local. Adversary callbacks are required to be global but the
+callbacks we construct are allocated on the heap and also global.
 
@@
 In summary, the use of local capabilities for other things than stack and return
-pointers is likely only possible in very specific scenario's when using our
+pointers is likely only possible in very specific scenarios when using our
 calling convention. While this is unfortunate, it is not unheard of that

### 2018-01-18 esop-2018/paper.tex
Fixed some of the todos with Lars. Citations and attacker model still on todo.

@@
 %% Title information
-\title{Reasoning About a Capability Machine with Local Capabilities}         %% [Short Title] is optional;
+%\title{Reasoning About a Capability Machine with Local Capabilities}         %% [Short Title] is optional;
+\title{Reasoning About a Machine with Local Capabilities}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Provably Safe Stack and Return Pointer Management (without OS Support)}                     %% \subtitle is optional
+%\subtitle{Provably Safe Stack and Return Pointer Management (without OS Support)}                     %% \subtitle is optional
+\subtitle{Provably Safe Stack and Return Pointer Management}                     %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
   Capability machines provide security guarantees at machine level which makes
-  them an interesting target for compilation schemes that provably enforce
-  properties like control-flow correctness and encapsulation of local state. We
-  provide a formalization of a representative capability machine with local
-  capabilities and study a novel calling convention for enforcing control-flow
-  correctness and encapsulation of local state on it. To prove these properties,
-  we provide a logical relation that semantically captures the guarantees
-  provided by the hardware (a form of capability safety). The logical relation
-  is not specific to our calling convention and can be used to reason about
-  arbitrary programs.
+  them an interesting target for secure compilation schemes that provably
+  enforce properties such as control-flow correctness and encapsulation of local
+  state. We provide a formalization of a representative capability machine with
+  local capabilities and study a novel calling convention.  We provide a logical
+  relation that semantically captures the guarantees provided by the hardware (a
+  form of capability safety) and use it to prove control-flow correctness and
+  encapsulation of local state.  The logical relation is not specific to our
+  calling convention and can be used to reason about arbitrary programs.
 \end{abstract}
@@
 the stack is not built-in, but merely an abstraction we put on this piece of the
-memory. The stack pointer is local and has $\rwlx$ permission. Crucially, the
-stack is the only part of memory for which the runtime (including malloc,
-loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
+memory. The stack pointer is local and has $\rwlx$ permission. Note that this
+means that we will be placing and executing instructions on the stack.
+Crucially, the stack is the only part of memory for which the runtime (including
+malloc, loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
 Additionally, our examples typically also assume some memory to store
@@
 
-\textbf{Reusable macro instructions} We define a number of reusable macros
+\textbf{Reusable macro instructions} 
+\label{para:macros}
+We define a number of reusable macros
 capturing the calling convention and other conveniences. All macros that use the
@@
 \label{sec:examples}
-In this section, we demonstrate how our formalization of capability
-safety allows us to prove local-state encapsulation and control-flow
-correctness properties for challenging program examples. The security
-measures of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to
-ensure these properties. Since we are dealing with assembly language,
-there are many details to the formal treatment, and therefore we
-necessarily omit some details in the lemma statements. The
-examples may look deceivingly short, but it is because they use the
-previously described macro instructions, which expands to multiple
-basic instructions. The examples would be unintelligible without the
-macros. The interested reader can find all the technical details in
-the technical appendix~\cite{technical_appendix}.
+In this section, we demonstrate how our formalization of capability safety
+allows us to prove local-state encapsulation and control-flow correctness
+properties for challenging program examples. The security measures
+of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to ensure these
+properties. Since we are dealing with assembly language, there are many details
+to the formal treatment, and therefore we necessarily omit some details in the
+lemma statements. The examples may look deceivingly short, but it is because
+they use the macro instructions described in Section \ref{para:macros}.
+The examples would be
+unintelligible without the macros, as each macro expands to multiple basic instructions. The interested reader can find all the
+technical details in the technical appendix~\cite{technical_appendix}.
 
@@
 
+\section*{Acknowledgements}
+This research was supported in part by the ModuRes Sapere Aude Advanced Grant
+from The Danish Council for Independent Research for the Natural Sciences (FNU).
+
 %% Acknowledgments

### 2018-01-18 esop-2018/paper.tex
added two sentences to section 3 pointing out modularity of CC (see todo)

@@
 scheme, we put local capabilities to work and take a number of
-not-always-obvious safety measures. In the next paragraphs, we will explain the
-issues to be considered in all the relevant situations: when (1) starting our
-program, (2) returning to the adversary, (3) invoking the adversary, (4)
-returning from the adversary, (5) invoking an adversary callback and (6) having
-a callback invoked by the adversary.
+not-always-obvious safety measures. The safety measures are presented in terms
+of what \emph{we} need to do to protect ourselves against an \emph{adversary},
+but this is for presentation purposes as our code assumes no special status on
+the machine. In fact, an adversary can apply the same safety measures to protect
+themselves agains us. In the next paragraphs, we will explain the issues to be
+considered in all the relevant situations: when (1) starting our program, (2)
+returning to the adversary, (3) invoking the adversary, (4) returning from the
+adversary, (5) invoking an adversary callback and (6) having a callback invoked
+by the adversary.
 

### 2018-01-18 esop-2018/paper.tex
Fixed citations through out the paper.

@@
 attacks (leading to an attack-defense arms race) and/or rely on techniques like
-machine code rewriting
-(e.g.,~\cite{wahbe_efficient_1993,abadi_control-flow_2005}), machine code
-verification (e.g.,~\cite{morrisett_system_1999}), virtual machines with a
-native stack (e.g.,~the JVM~\cite{lindholm_java_2014}) or
+machine code rewriting~\cite{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
+verification~\cite{morrisett_system_1999}, virtual machines with a
+native stack~\cite{lindholm_java_2014} or
 randomization~\cite{forrest_building_1997}. The latter techniques essentially
@@
 remediate these limitations with a better security model at
-the hardware level. They are based on old ideas
-(e.g.~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999}),
+the hardware level. They are based on old ideas~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
 but have recently received renewed interest; in particular, the CHERI
@@
   \emph{single} orthogonal closure (rather than the earlier used biorthogonal
-  closure) and a variant of Dreyer et. al.'s public and private future
+  closure) and a variant of Dreyer et al.~\cite{Dreyer:jfp12}'s public and private future
   worlds~\cite{Dreyer:jfp12} to express the special nature of local
@@
 requirements on the memory. It simply serves as a mask for a revoked temporary
-region. Masking a region like this goes back to earlier work of
-\cite{Ahmed2004semantics} and was also used by
-\cite{Thamsborg:2011:KLR:2034773.2034831}.
+region. Masking a region like this goes back to earlier work of Ahmed
+\cite{Ahmed2004semantics} and was also used by Birkedal et al.~\cite{Thamsborg:2011:KLR:2034773.2034831}.
 
@@
 in Kripke models and the following lemma uses the method of
-\cite{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
+Birkedal and Bizjak~\cite{Birkedal:2011:SKM:1926385.1926401}; Birkedal et al.~\cite{Birkedal:tutorial-notes} for
 constructing them. The formulation of the lemma is technical, so we recommend
@@
 In fact, the machine often resorts to failing when an unauthorized access is
-attempted, like loading from a capability without read permission. This is
-similar to \cite{Devriese:2016ObjCap}'s logical relation for an untyped
+attempted, such as loading from a capability without read permission. This is
+similar to Devriese et al.~\cite{Devriese:2016ObjCap}'s logical relation for an untyped
 language, but unlike typical logical relations for typed languages, which
@@
 Formulations in terms of a control-flow graph, like the one by
-\cite{abadi_control-flow_2005}, do not take into account temporal properties,
+Abadi et al.~\cite{abadi_control-flow_2005}, do not take into account temporal properties,
 like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
@@
 
-Interestingly, \cite{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
+Interestingly, Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
 Kripke logical relation for an assembly language (for reasoning about correct
@@
 code that relies on this (particularly, the original awkward example),
-\cite{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
+Dreyer et al.~\cite{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
 contexts using Kripke worlds with public and private future world relations.
@@
 local capabilities are quite similar to a feature proposed for the high-level
-language Scala: \cite{osvald_gentrification_2016}'s second-class or local
+language Scala: Osvald et al.~\cite{osvald_gentrification_2016}'s second-class or local
 values. They are a kind of values that can be provided to other code for
@@
 
-Capability machines originate with \cite{Dennis:1966:PSM:365230.365252} and we refer to
-\cite{Levy1984capability} and \cite{Watson2015Cheri} for an overview
+Capability machines originate with Dennis and Van Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
+Levy~\cite{Levy1984capability} and Watson et al.~\cite{Watson2015Cheri} for an overview
 of previous work. The capability machine formalized in
@@
 \cite{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
-closely related to the one used by \cite{Devriese:2016ObjCap} to formulate
+closely related to the one used by Devriese et al.~\cite{Devriese:2016ObjCap} to formulate
 capability safety in a high-level JavaScript-like lambda calculus. Our
@@
 
-Swasey et. al.~\cite{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
+Swasey et al.~\cite{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
 of object capability patterns. The logic is based on
@@
 capability safety for a lambda-calculus (with concurrency) compared to the
-earlier work by \cite{Devriese:2016ObjCap}.
+earlier work by Devriese et al.~\cite{Devriese:2016ObjCap}.
 %domi: rest of the paragraph could be cut?

### 2018-01-18 esop-2018/paper.tex
changed acks to subsubsection

@@
 
-\section*{Acknowledgements}
+\subsubsection*{Acknowledgements.}
 This research was supported in part by the ModuRes Sapere Aude Advanced Grant

### 2018-01-18 esop-2018/paper.tex
mend

@@
 of what \emph{we} need to do to protect ourselves against an \emph{adversary},
-but this is for presentation purposes as our code assumes no special status on
+but this is only for presentation purposes as our code assumes no special status on
 the machine. In fact, an adversary can apply the same safety measures to protect
-themselves agains us. In the next paragraphs, we will explain the issues to be
+themselves against us. In the next paragraphs, we will explain the issues to be
 considered in all the relevant situations: when (1) starting our program, (2)

### 2018-01-18 fully-abstract-comp/technical-report.tex
a couple of typos in the lrv.
a bit of formatting to make overflowing part fit on page
Some minor comments and typo fixes as I was going through the TR.

@@
     &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
-    &\scbnf &\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
+    \scbnf \in &\SealableCaps&\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    &\cbnf & \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
+    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
     &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
@@
 Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}
+\dominique{18-1-2018: Lau, I suppose the above has been answered verbally already during one of the previous skype calls, but for reference, this is what the spatial vs spatial_owned distinction is intended to solve: the world will contain spatial regions for such memory segments, but not spatial_owned ones, so the memory satisfaction will require the memory to be present but not allow it to be addressed.}
+\dominique{18-1-2018: $W_\lrk$, $k_{d,*}$, $k_{c,*}$ can all be removed from the above definition now, right?}
 
@@
 \]
-Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
+% DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
+% Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
+\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
 
@@
       \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
-      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}{W_\lrm}} \wedge \\
+      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm}} \wedge \\
       \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata,\rstk,\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\

### 2018-01-19 fully-abstract-comp/technical-report.tex
Replied to Dominiques comments (and made a few corrections based on them).

@@
 \subsection{Future world}
-
 For both $\privft$ and $\pubft$ we have
@@
     \begin{array}{l}
-      \forall n' \leq n, \src{k_{d,S}}, \src{k_{c,S}}, k_{d,T}, k_{c,T}, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \forall W_\lrk, W_\lrr, W_\lrm \ldotp \\
+      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad \forall W_\lrr, W_\lrm \ldotp \\
       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_\lrr) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
       \qquad\Rightarrow\\
-      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,S},v_{d,S},k_{c,S},k_{d,S}},\stk, \ms_\stk)},\\
-                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata,\rretc,\rretd}{v_{c,T},v_{d
-,T},k_{c,T},k_{d,T}})\endarray\right)}
-      \in \lro(W\oplus W_\lrk \oplus W_\lrr \oplus W_\lrm)
+      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata}{v_{c,S},v_{d,S}},\stk, \ms_\stk)},\\
+                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata}{v_{c,T},v_{d,T}})
+                       \endarray\right)}
+      \in \lro(W \oplus W_\lrr \oplus W_\lrm)
     \end{array}
@@
 \lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out.
-I can see how it will enforce linearity, but is you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part.
+I can see how it will enforce linearity, but if you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part.
 In other words, the memory satisfaction will not be able to find a concrete memory segment for segments governed by a linear capability in the register file.
 Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}
-\dominique{18-1-2018: Lau, I suppose the above has been answered verbally already during one of the previous skype calls, but for reference, this is what the spatial vs spatial_owned distinction is intended to solve: the world will contain spatial regions for such memory segments, but not spatial_owned ones, so the memory satisfaction will require the memory to be present but not allow it to be addressed.}
+\dominique{18-1-2018: Lau, I suppose the above has been answered verbally already during one of the previous skype calls, but for reference, this is what the spatial vs spatial\_owned distinction is intended to solve: the world will contain spatial regions for such memory segments, but not spatial\_owned ones, so the memory satisfaction will require the memory to be present but not allow it to be addressed.}
 \dominique{18-1-2018: $W_\lrk$, $k_{d,*}$, $k_{c,*}$ can all be removed from the above definition now, right?}
-
+\lau{19-01-2018: Yes. And I have removed them}
 \[
@@
       \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
-      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg \} \cup R))} S(r) \\
+      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \\
       \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
@@
 \]
+We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simlpy omit that argument.
 % DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
 % Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
-\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
+%\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
+% Lau 19-01-2018: Added rdata as default.
 
@@
       (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
-      \npair{(\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})} \in \lrp(W)
+      \npair{([\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})]} \in \lrp(W)
     \end{array}
@@
       \forall n' \leq n \ldotp \forall W' \privft W \ldotp \\
-      \quad \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
+      \quad \npair[n']{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
       \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
-      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm}} \wedge \\
-      \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata,\rstk,\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
+      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
       \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
@@
       \qquad \Rightarrow \\
-      \qquad \quad \npair[n']{\left(\array{l}(\ms_S,\reg_S\updReg{\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}}{\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0},\stk,\ms_\stk \uplus \ms_f),\\
-                                             (\ms_T,\reg_S\updReg{\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}}{((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?)})
+      \qquad \quad \npair[n']{\left(\arraycolsep=0pt\array{l}
+                                     \left(\arraycolsep=0pt\array{l}\ms_S,\reg_S \left[
+                                       \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0\endarray \right]\\
+                                     ,\stk,\ms_\stk \uplus \ms_f
+                                     \endarray \right),\\
+                                             \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0\endarray\right]\right)
                                     \endarray\right)} \in \lro(W_\var{new})

### 2018-01-19 fully-abstract-comp/technical-report.tex
A quick and dirty write-up of possible new future world relation.

@@
 \subsection{Future world}
+\subsubsection{\color{green} first thoughts on change to future world relation}
+{\color{green}
+The public part of the future world relation stays the same:
+\begin{mathpar}
+  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \pubft W(r) }
+            { W' \future W }
+\end{mathpar}
+The private future world relation becomes:
+\begin{mathpar}
+  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \privft W(r) \\  \exists r \in \dom(W') \cap \dom(W)\ldotp W'(r) \text{ is non-empty and stack region}}
+            { W' \future W }
+\end{mathpar}
+where we take "is non-empty and stack region" to mean the following:
+
+\begin{definition}
+  We say a region $(k,s,\phi_\pub,\phi,H)$ is non-empty iff
+\[
+  \begin{array}{l}
+    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) > 0 \wedge \dom(\ms_T) > 0 
+  \end{array}
+\]
+\end{definition}
+
+\begin{definition}
+  We say a region $(k,s,\phi_\pub,\phi,H)$ is a stack region on a machine with starting stack domain $[\stkb,\eaddr]$ iff
+\[
+  \begin{array}{l}
+    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) \subset [\stkb,\eaddr]  \wedge \dom(\ms_T) \subset [\stkb,\eaddr]
+  \end{array}
+\]
+\end{definition}
+
+}
+
 For both $\privft$ and $\pubft$ we have

### 2018-01-19 fully-abstract-comp/technical-report.tex
typos in proposed ft world relation

@@
 \begin{mathpar}
-  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \privft W(r) \\  \exists r \in \dom(W') \cap \dom(W)\ldotp W'(r) \text{ is non-empty and stack region}}
+  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \privft W(r) \\  \exists r \in \dom(W') \setminus \dom(m(W))\ldotp W'(r) \text{ is non-empty and stack region}}
             { W' \future W }
@@
   \begin{array}{l}
-    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) \subset [\stkb,\eaddr]  \wedge \dom(\ms_T) \subset [\stkb,\eaddr]
+    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) \subseteq [\stkb,\eaddr]  \wedge \dom(\ms_T) \subseteq [\stkb,\eaddr]
   \end{array}
@@
 \end{definition}
-
 }

### 2018-01-22 fully-abstract-comp/technical-report.tex
A few corrections
Propose refactoring to merge cases for return and closure code-data-pairs.

@@
 \newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
+\newcommand{\xjumpResult}[3]{\plainfun{xjumpResult}{#1,#2,#3}}
 
@@
 
+\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
+\begin{align*}
+  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
+                                   \begin{cases}
+                                       \Phi''
+                                     &
+                                     \begin{array}[t]{l}
+                                       \Phi(r_1) = c_1 \tand
+                                       \Phi(r_2) = c_2 \tand\\
+                                       w_1 = \linCons{c_1} \tand \\
+                                       w_2 = \linCons{c_2} \tand \\
+                                       \Phi' = \Phi \updReg{r_1}{w_1} \updReg{r_2}{w_2} \tand\\
+                                       \Phi'' = \xjumpResult{c_1}{c_2}{\Phi'}
+                                     \end{array}\\
+                                     \failed & \totherwise
+                                   \end{cases}
+\end{align*}
+\begin{align*}
+  \xjumpResult{c_1}{c_2}{\Phi} = 
+  \begin{cases}
+    \arraycolsep=0pt
+    \begin{array}[t]{rl}
+      \Phi & \updReg{\pcreg}{c_\code}\\
+           & \updReg{\rdata}{c_\data}
+    \end{array} &
+    \begin{array}[t]{l}
+      c_1 = \sealed{\sigma_1,c_\code} \tand \\
+      c_2 = \sealed{\sigma_2,c_\data} \tand \\
+      \sigma_1 = \sigma_2 \tand \\
+      \nonExec{c_\data}
+    \end{array}
+    \\&\\
+    \sourcecolor\left.
+      \arraycolsep=0pt
+      \begin{array}[t]{rl}
+        \Phi' & \updReg{r_2}{0} \\
+              & \updReg{\pcreg}{\opc}\\
+              & \updReg{\rdata}{0}\\
+              & \updReg{\rstk}{c_\stk}\\
+              & \updReg{\rtmp{1}}{0}\\
+              & \updReg{\rtmp{2}}{0}
+      \end{array}\right.&
+    \sourcecolor\left.
+      \begin{array}[t]{l}
+        c_1 = \retptrc(\sigma_1,\_) \tand \\
+        c_2 = \retptrd(\sigma_2,(\aaddr_\stk,\_)) \tand \\
+        \sigma_1 = \sigma_2 \tand\\
+        \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
+        \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
+        \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
+        \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
+        c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
+        \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
+      \end{array}\right.
+    \\
+    \tundefined & \totherwise
+  \end{cases}
+\end{align*}
+
 \noindent\textbf{cseal}\\
@@
 \dominique{18-1-2018: Lau, I suppose the above has been answered verbally already during one of the previous skype calls, but for reference, this is what the spatial vs spatial\_owned distinction is intended to solve: the world will contain spatial regions for such memory segments, but not spatial\_owned ones, so the memory satisfaction will require the memory to be present but not allow it to be addressed.}
-\dominique{18-1-2018: $W_\lrk$, $k_{d,*}$, $k_{c,*}$ can all be removed from the above definition now, right?}
-\lau{19-01-2018: Yes. And I have removed them}
+
+\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
+\[
+  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+    \begin{array}{l}
+      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad \forall W_\lrr, W_\lrm \ldotp \\
+      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
+      \qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S}\\
+      \qquad\Phi_T = (\ms_T,\reg_T)\\
+      \qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro(W \oplus W_\lrr \oplus W_\lrm)
+    \end{array}
+    \right\}
+\]
+
 \[
@@
 \lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}
+
+\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
+\[
+  \lrv(H_\sigma,W) =
+  \begin{array}[t]{l}
+    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{c_S},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
+    \begin{array}{l}
+      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
+      \forall W' \future W, W_o, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
+      \quad \npair{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lre(W'\oplus W_o) \wedge \\
+      \quad \npair{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lre(W' \oplus W_o)
+    \end{array}
+    \right\}\cup\\
+%
+    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    % \right\} \cup \\
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \begin{array}{r l l}
+      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
+      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
+    \end{array}
+    \right\} \cup \\
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \begin{array}{r l l }
+      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
+      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
+      \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
+      \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
+    \end{array}
+    \right\}
+  \end{array}
+\]
 

### 2018-01-23 fully-abstract-comp/technical-report.tex
Fix my previous refactoring proposal.
In the process, this updated refactoring simplifies a number of things by
splitting the expression relation into two parts: one for pairs of caps being
xjumped to and one for regular caps being regularly jumped to.

@@
 \newcommand{\lre}{\typesetlr{E}}
+\newcommand{\lrexj}{\typesetlr{E}_{\var{xjmp}}}
 \newcommand{\lrk}{\typesetlr{K}}
@@
 It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
-\[
-  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
+The argument for having one relation relate pairs of pairs of capabilties and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
+\begin{align*}
+  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
     \begin{array}{l}
@@
     \end{array}
+    \right\}\\
+  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
+    \begin{array}{l}
+      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad \forall W_\lrr, W_\lrm \ldotp \\
+      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}}]} \in \lrp(W) \wedge \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
+      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
+      \qquad\Phi_T = (\ms_T,\reg_T)\\
+      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro(W \oplus W_\lrr \oplus W_\lrm)
+    \end{array}
     \right\}
-\]
+\end{align*}
 
@@
       \forall W' \future W, W_o, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
-      \quad \npair{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lre(W'\oplus W_o) \wedge \\
-      \quad \npair{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lre(W' \oplus W_o)
+      \quad \npair{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
+      \quad \npair{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
     \end{array}
@@
       \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
+    \end{array}
+    \right\}
+\]
+
+\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
+Note: this new version of execCond uses the expression relation for regular jumps since it expresses the validity of jumping to an address pointed to by an executable capability. 
+\[
+  \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
+    \begin{array}{l}
+      \forall n' < n, W' \future W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr], W'' \ldotp\\
+      \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}]} \in \lrp(W' \oplus W'') \Rightarrow \\
+      \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}} \in \lre(W' \oplus W'')
     \end{array}
@@
                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray} \in \lre(W)
+
+    \end{array}
+  \]
+\end{theorem}
+
+\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
+Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps.
+It expresses that any reasonable executable capability is safe to use as the pc capability.
+\dominique{I'm following the example of the above FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
+\begin{theorem}[FTLR]
+  \label{thm:ftlr}
+  \[
+    \begin{array}{l}
+      \forall n,W,\perm,\lin,\baddr,\eaddr,\aaddr \ldotp \\
+      \quad \left(
+      \begin{array}{l}
+        \perm_C = \rx \wedge\\
+        \npair{(\baddr_C,\eaddr_C)} \in \readCond{\lin_C,W_C}
+      \end{array}
+      \right)\\
+      \quad\Rightarrow 
+      \npair{\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),\\
+                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray} \in \lre(W)
 

### 2018-01-25 fully-abstract-comp/technical-report.tex
A couple of technicalities I suspect are necessary.

@@
 \documentclass[a3paper]{article}
-\usepackage[a3paper]{geometry}
+\usepackage[a3paper,margin=0.7in]{geometry}
 \usepackage[utf8]{inputenc}
@@
     \text{// clear temporary registers used}\\
-    \tmove{\rtmp{1}}{0}\\
     \tmove{\rtmp{2}}{0}\\
@@
   \item adjust stack pointer to first empty address (the address with the end address is considered free)  
-  \item clear the temporary registers.
+  \item clear the temporary registers. (Note that $\rtmp{1}$ is not cleared with 0 as it already contains 0 after the $\stkb$ check. If it didn't contain 0, then the execution would have failed.)
+
   \end{itemize}
@@
                                          \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
-                                         \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\_)) \tand \\
+                                         \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
                                          \sigma_1 = \sigma_2 \tand\\
@@
                                          \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
+                                         \eaddr_{\stk} + 1 = \aaddr_\stk \tand\\
                                          c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
@@
       \begin{array}[t]{rl}
-        \Phi' & \updReg{r_2}{0} \\
+        \Phi' & \updReg{r_1}{w_1}\\
+              & \updReg{r_2}{0} \\
               & \updReg{\pcreg}{\opc}\\
@@
       \begin{array}[t]{l}
+        w_1 = \linCons{c_1}
         c_1 = \retptrc(\sigma_1,\_) \tand \\
-        c_2 = \retptrd(\sigma_2,(\aaddr_\stk,\_)) \tand \\
+        c_2 = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
         \sigma_1 = \sigma_2 \tand\\
@@
         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
+        e_\stk + 1 = a_\stk \tand\\
         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
@@
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge\\
+      \qquad W_{\var{new}} = W \oplus W_\lrr \oplus W_\lrm \\
       \qquad\Rightarrow\\
@@
                        \endarray\right)}
-      \in \lro(W \oplus W_\lrr \oplus W_\lrm)
+      \in \lro(W_{\var{new}})
     \end{array}
@@
       \qquad \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; S(i)(r).s \; \xi^{-1}(S(i))  \ldotp \\
-      \qquad \qquad \dom(\src{\ms_S}) \subseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \subseteq [\baddr,\eaddr]      
+      \qquad \qquad \dom(\src{\ms_S}) \supseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \supseteq [\baddr,\eaddr]      
     \end{array}
@@
       \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
-      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \\
+      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \wedge \\
       \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\\sealed{\sigma,\vsc_C}\endarray\right)} \;\middle| \;
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\
+                         \sealed{\sigma,((\perm,\lin),\baddr,\eaddr,\aaddr + ?)}\endarray\right)} \;\middle| \;
     \begin{array}{l}
@@
       \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
-      \forall n' \leq n \ldotp \forall W' \privft W \ldotp \\
+      \forall n' \leq n \ldotp \forall W' \privft W , W_C,W_\lrm,W_\lrr\ldotp \\
       \quad \npair[n']{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
-      \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
+      \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\
+                                                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
       \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \quad \opc = ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C) \wedge \\
       \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
@@
       \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
-      \quad W_\var{new} = (W' \uplus W_\lrr \uplus W_\lrm \uplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
+      \quad W_\var{new} = (W' \oplus W_\lrr \oplus W_\lrm \oplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
       \qquad \Rightarrow \\
@@
                                      \left(\arraycolsep=0pt\array{l}\ms_S,\reg_S \left[
-                                       \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0\endarray \right]\\
+      \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0,0\endarray \right]\\
                                      ,\stk,\ms_\stk \uplus \ms_f
                                      \endarray \right),\\
-                                             \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2}\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0\endarray\right]\right)
+                                             \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0,0\endarray\right]\right)
                                     \endarray\right)} \in \lro(W_\var{new})
@@
       \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
-      \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
-      \quad \npair{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lre(W) \wedge \\
-      \quad \npair{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lre(W)
+      \forall W' \pubft W, n' \leq n \ldotp \\
+      \quad \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
+      \qquad \npair[n']{(\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T')} \in \lre(W') \wedge \\
+      \qquad \npair[n']{(\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T)} \in \lre(W')
     \end{array}
@@
     \begin{array}{l}
+      \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
       \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
-      \forall W' \future W, W_o, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
-      \quad \npair{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
-      \quad \npair{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
+      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
+      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
+      \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
     \end{array}
@@
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \begin{array}{l}
+      \npair{(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
       \begin{array}{r l l}
-      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W} \wedge \\
         \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
+      \end{array}
     \end{array}
@@
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \begin{array}{l}
+      \npair{(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
       \begin{array}{r l l }
@@
         \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
+      \end{array}
     \end{array}
@@
       \quad\Rightarrow 
-      \npair{\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
-                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray} \in \lre(W)
+      \npair{\left(\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
+                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray \right)} \in \lre(W)
 
@@
     \begin{array}{l}
-      \forall n,W,\perm,\lin,\baddr,\eaddr,\aaddr \ldotp \\
+      \forall n,W,\lin,\baddr,\eaddr,\aaddr \ldotp \\
       \quad \left(
       \begin{array}{l}
-        \perm_C = \rx \wedge\\
-        \npair{(\baddr_C,\eaddr_C)} \in \readCond{\lin_C,W_C}
+        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}
       \end{array}
@@
       \quad\Rightarrow 
-      \npair{\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),\\
-                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray} \in \lre(W)
+      \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
+                      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)
 

### 2018-01-25 fully-abstract-comp/technical-report.tex
put the preamble in its own file

@@
 
-% Math packages
-\usepackage{amsmath,amsfonts,amssymb,amsthm}
-\usepackage{mathrsfs}
-\usepackage{thmtools}
-\usepackage{array}
-\usepackage{cleveref}
-\usepackage{stmaryrd}
-\usepackage{mathpartir}
-
 % References
@@
 
-% Command control packages
-\usepackage{ifthen}
-\usepackage{ifpdf}
-
-% Listings
-\usepackage{listings}
-\lstset{
-  basicstyle=\ttfamily,
-  columns=fullflexible,
-  keepspaces=true,
-  mathescape
-}
-
 % Rotate entire figures
@@
 
-% Tikz
-\usepackage{tikz}
-
-%%% Comments
-% Comments
-\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
-\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
-\newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
-
-%%% Math environments
-\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
-\declaretheorem[numbered=yes,name=Theorem,qed=$\blacksquare$]{theorem}
-\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
-\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}
-
-%%% Math notation
-\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
-\newcommand{\defbnf}{::=}
-\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
-\newcommand{\ssem}[2][\Phi]{\sem{#2}_{\mathrm{src}}(#1)}
-\newcommand{\tsem}[2][\Phi]{\sem{#2}_{\mathrm{trg}}(#1)}
-\newcommand{\dom}{\mathrm{dom}}
-\newcommand{\powerset}[1]{\mathcal{P}(#1)}
-
-\newcommand{\npair}[2][n]{\left(#1,#2\right)}
-
-\newcommand{\nsubeq}[1][n]{\overset{#1}{\subseteq}}
-\newcommand{\nsupeq}[1][n]{\overset{#1}{\supseteq}}
-\newcommand{\nequal}[1][n]{\overset{#1}{=}}
-
-% Function arrows
-\newcommand{\fun}{\rightarrow}
-\newcommand{\parfun}{\rightharpoonup}
-\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
-
-% Text
-\newcommand{\tand}{\text{ and }}
-\newcommand{\tor}{\text{ or }}
-\newcommand{\totherwise}{\text{otherwise }}
-
-% Equivalences
-\newcommand{\sconeq}{\mathrel{\src{\approx_{\mathrm{ctx}}}}}
-\newcommand{\tconeq}{\mathrel{\approx_{\mathrm{ctx}}}}
-
-%%% Logical Relation notation
-\newcommand{\typesetlr}[1]{\mathcal{#1}}
-\newcommand{\lre}{\typesetlr{E}}
-\newcommand{\lrexj}{\typesetlr{E}_{\var{xjmp}}}
-\newcommand{\lrk}{\typesetlr{K}}
-\newcommand{\lrr}{\typesetlr{R}}
-\newcommand{\lro}{\typesetlr{O}}
-\newcommand{\lrv}{\typesetlr{V}}
-\newcommand{\lrp}{\typesetlr{P}}
-\newcommand{\lrm}{\typesetlr{M}}
-
-\newcommand{\stpair}[3][]{
-\ifthenelse{\equal{#1}{}}
-{\left(\src{#2_S},#3_T\right)}
-{\left(\src{#2},#3\right)}}
-
-\newcommand{\memSat}[3][n]{#2 :_{#1}#3}
-
-\newcommand{\World}{\mathrm{World}}
-\newcommand{\RegionName}{\mathrm{RegionName}}
-\newcommand{\Region}{\mathrm{Region}}
-\newcommand{\spatial}{\mathrm{spatial}}
-\newcommand{\spatialo}{\mathrm{spatial\_owned}}
-\newcommand{\pure}{\mathrm{pure}}
-\newcommand{\revoked}{\mathrm{revoked}}
-\newcommand{\State}{\mathrm{State}}
-\newcommand{\Rels}{\mathrm{Rels}}
-\newcommand{\UPred}[1]{\mathrm(#1)}
-
-\newcommand{\future}{\sqsupseteq}
-\newcommand{\pub}{\mathrm{pub}}
-\newcommand{\privft}{\future^{\priv}}
-\newcommand{\pubft}{\future^{\pub}}
-\newcommand{\monprivnefun}{\xrightarrow[\text{\tiny{$\privft$}}]{\textit{\tiny{mon, ne}}}}
-\newcommand{\monpubnefun}{\xrightarrow[\text{\tiny{$\pubft$}}]{\textit{\tiny{mon, ne}}}}
-
-%%% Regions
-\newcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
-\newcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
-\newcommand{\spa}{\mathrm{s}}
-\newcommand{\spao}{\mathrm{so}}
-\newcommand{\pur}{\mathrm{p}}
-
-%%% Instruction formatting
-\newcommand{\sourcecolortext}{blue}
-\newcommand{\sourcecolor}{\color{blue}}
-\newcommand{\src}[1]{{\sourcecolor #1}}
-\newcommand{\targetcolortext}{black}
-\newcommand{\targetcolor}[1]{\color{black}}
-\newcommand{\trg}[1]{{\targetcolor{} #1}}
-
-\newcommand{\zinstr}[1]{\texttt{#1}}
-\newcommand{\oneinstr}[2]{
-  \ifthenelse{\equal{#2}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2}
-}
-\newcommand{\twoinstr}[3]{
-  \ifthenelse{\equal{#2#3}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3}
-}
-\newcommand{\threeinstr}[4]{
-  \ifthenelse{\equal{#2#3#4}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3 \; #4}
-}
-
-\newcommand{\fourinstr}[5]{
-  \ifthenelse{\equal{#2#3#4#5}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3 \; #4 \; #5}
-}
-
-%%% Source language
-% No arguments
-\newcommand{\sfail}{\zinstr{\src{fail}}}
-\newcommand{\shalt}{\zinstr{\src{halt}}}
-\newcommand{\sreturn}{\zinstr{\src{return}}}
-
-% One argument
-\newcommand{\sjmp}[1]{\oneinstr{\src{jmp}}{#1}}
-\newcommand{\spush}[1]{\oneinstr{\src{push}}{#1}}
-\newcommand{\spop}[1]{\oneinstr{\src{pop}}{#1}}
-
-% Two arguments
-\newcommand{\sjnz}[2]{\twoinstr{\src{jnz}}{#1}{#2}}
-\newcommand{\sisptr}[2]{\twoinstr{\src{gettype}}{#1}{#2}}
-\newcommand{\sgeta}[2]{\twoinstr{\src{geta}}{#1}{#2}}
-\newcommand{\sgetb}[2]{\twoinstr{\src{getb}}{#1}{#2}}
-\newcommand{\sgete}[2]{\twoinstr{\src{gete}}{#1}{#2}}
-\newcommand{\sgetp}[2]{\twoinstr{\src{getp}}{#1}{#2}}
-%\newcommand{\sgetloc}[2]{\twoinstr{\src{getloc}}{#1}{#2}}
-\newcommand{\sgetlin}[2]{\twoinstr{\src{get}}{#1}{#2}}
-\newcommand{\smove}[2]{\twoinstr{\src{move}}{#1}{#2}}
-\newcommand{\sstore}[2]{\twoinstr{\src{store}}{#1}{#2}}
-\newcommand{\sload}[2]{\twoinstr{\src{load}}{#1}{#2}}
-\newcommand{\scca}[2]{\twoinstr{\src{cca}}{#1}{#2}}
-\newcommand{\ssload}[2]{\twoinstr{\src{sload}}{#1}{#2}}
-\newcommand{\sxjmp}[2]{\twoinstr{\src{xjmp}}{#1}{#2}}
-\newcommand{\ssetatob}[2]{\twoinstr{\src{seta2b}}{#1}{#2}}
-% scall - special two instruction
-\newcommand{\scall}[4][]{  
-\ifthenelse{\equal{#3#4}{}}
-  {\ensuremath{\zinstr{\src{call}}_{#1}^{#2}}}
-  {\ensuremath{\zinstr{\src{call}}_{#1}^{#2} \; #3 \; #4}}
-}
-
-% Three arguments
-\newcommand{\srestrict}[3]{\threeinstr{\src{restrict}}{#1}{#2}{#3}}
-\newcommand{\slt}[3]{\threeinstr{\src{lt}}{#1}{#2}{#3}}
-\newcommand{\splus}[3]{\threeinstr{\src{plus}}{#1}{#2}{#3}}
-\newcommand{\sminus}[3]{\threeinstr{\src{minus}}{#1}{#2}{#3}}
-\newcommand{\scseal}[3]{\threeinstr{\src{cseal}}{#1}{#2}{#3}}
-\newcommand{\ssplice}[3]{\threeinstr{\src{splice}}{#1}{#2}{#3}}
-
-% Four arguments
-%\newcommand{\ssubseg}[4]{\fourinstr{\src{subseg}}{#1}{#2}{#3}{#4}}
-\newcommand{\ssplit}[4]{\fourinstr{\src{split}}{#1}{#2}{#3}{#4}}
-
-%%% Target language
-% No arguments
-\newcommand{\tfail}{\zinstr{\trg{fail}}}
-\newcommand{\thalt}{\zinstr{\trg{halt}}}
-
-% One argument
-\newcommand{\tjmp}[1]{\oneinstr{\trg{jmp}}{#1}}
-
-% Two arguments
-\newcommand{\tjnz}[2]{\twoinstr{\trg{jnz}}{#1}{#2}}
-\newcommand{\tisptr}[2]{\twoinstr{\trg{gettype}}{#1}{#2}}
-\newcommand{\tgeta}[2]{\twoinstr{\trg{geta}}{#1}{#2}}
-\newcommand{\tgetb}[2]{\twoinstr{\trg{getb}}{#1}{#2}}
-\newcommand{\tgete}[2]{\twoinstr{\trg{gete}}{#1}{#2}}
-\newcommand{\tgetp}[2]{\twoinstr{\trg{getp}}{#1}{#2}}
-%\newcommand{\tgetloc}[2]{\twoinstr{\trg{getloc}}{#1}{#2}}
-\newcommand{\tgetlin}[2]{\twoinstr{\trg{getl}}{#1}{#2}}
-\newcommand{\tmove}[2]{\twoinstr{\trg{move}}{#1}{#2}}
-\newcommand{\tstore}[2]{\twoinstr{\trg{store}}{#1}{#2}}
-\newcommand{\tload}[2]{\twoinstr{\trg{load}}{#1}{#2}}
-\newcommand{\tcca}[2]{\twoinstr{\trg{cca}}{#1}{#2}}
-\newcommand{\txjmp}[2]{\twoinstr{\trg{xjmp}}{#1}{#2}}
-\newcommand{\tsetatob}[2]{\twoinstr{\trg{seta2b}}{#1}{#2}}
-
-% Three arguments
-\newcommand{\tsplice}[3]{\threeinstr{\trg{splice}}{#1}{#2}{#3}}
-\newcommand{\trestrict}[3]{\threeinstr{\trg{restrict}}{#1}{#2}{#3}}
-\newcommand{\tlt}[3]{\threeinstr{\trg{lt}}{#1}{#2}{#3}}
-\newcommand{\tplus}[3]{\threeinstr{\trg{plus}}{#1}{#2}{#3}}
-\newcommand{\tminus}[3]{\threeinstr{\trg{minus}}{#1}{#2}{#3}}
-\newcommand{\tcseal}[3]{\threeinstr{\trg{cseal}}{#1}{#2}{#3}}
-
-% Four arguments
-%\newcommand{\tsubseg}[4]{\fourinstr{\trg{subseg}}{#1}{#2}{#3}{#4}}
-\newcommand{\tsplit}[4]{\fourinstr{\trg{split}}{#1}{#2}{#3}{#4}}
-
-%%% Domains
-\newcommand{\plaindom}[1]{\mathrm{#1}}
-
-\newcommand{\nats}{\mathbb{N}}
-\newcommand{\ints}{\mathbb{Z}}
-
-%%% Updates
-\newcommand{\update}[2]{[#1 \mapsto #2]}
-\newcommand{\updReg}[2]{\update{\reg.#1}{#2}}
-\newcommand{\updPc}[1]{\Phi\updReg{\pcreg}{#1}}
-
-%%% Source dom
-\newcommand{\shareddom}[1]{\mathrm{#1}}
-\newcommand{\RegName}{\shareddom{RegisterName}}
-\newcommand{\Addr}{\shareddom{Addr}}
-\newcommand{\Seal}{\shareddom{Seal}}
-\newcommand{\Perm}{\shareddom{Perm}}
-\newcommand{\Caps}{\shareddom{Cap}}
-\newcommand{\SealableCaps}{\shareddom{SealableCap}}
-\newcommand{\Word}{\shareddom{Word}}
-\newcommand{\Instr}{\shareddom{Instr}}
-\newcommand{\Mem}{\shareddom{Memory}}
-\newcommand{\Reg}{\shareddom{RegisterFile}}
-\newcommand{\Stk}{\shareddom{Stack}}
-\newcommand{\Conf}{\shareddom{Conf}}
-\newcommand{\ExecConf}{\shareddom{ExecConf}}
-%\newcommand{\Global}{\shareddom{Global}}
-\newcommand{\Linear}{\shareddom{Linear}}
-\newcommand{\MemSeg}{\shareddom{MemorySegment}}
-\newcommand{\StkFrame}{\shareddom{StackFrame}}
-\newcommand{\Stack}{\shareddom{Stack}}
-
-\newcommand{\scbnf}{\shareddom{sc}}
-\newcommand{\cbnf}{\shareddom{c}}
-\newcommand{\permbnf}{\shareddom{perm}}
-\newcommand{\addrbnf}{\shareddom{a}}
-\newcommand{\basebnf}{\shareddom{base}}
-\newcommand{\aendbnf}{\shareddom{end}}
-\newcommand{\rbnf}{\shareddom{r}}
-%\newcommand{\glbnf}{\shareddom{gl}}
-\newcommand{\linbnf}{\shareddom{l}}
-\newcommand{\sealbasebnf}{\sigma_\shareddom{base}}
-\newcommand{\sealendbnf}{\sigma_\shareddom{end}}
-
-\newcommand{\sstk}{\shareddom{stk}}
-\newcommand{\smsstk}{\shareddom{ms_{stk}}}
-\newcommand{\sstkframe}{\shareddom{frame}}
-\newcommand{\sopc}{\shareddom{opc}}
-\newcommand{\sastk}{\shareddom{a_{stk}}}
-\newcommand{\perm}{\var{perm}}
-%\newcommand{\gl}{\var{g}}
-\newcommand{\lin}{\var{l}}
-\newcommand{\base}{\shareddom{base}}
-\newcommand{\aend}{\shareddom{end}}
-\newcommand{\addr}{\shareddom{a}}
-\newcommand{\scap}{\shareddom{c}}
-\newcommand{\sms}{\shareddom{ms}}
-
-\newcommand{\stkptr}[1]{\mathrm{stack\text{-}ptr}(#1)}
-\newcommand{\retptr}{\mathrm{ret\text{-}ptr}}
-\newcommand{\retptrd}{\mathrm{ret\text{-}ptr\text{-}data}}
-\newcommand{\retptrc}{\mathrm{ret\text{-}ptr\text{-}code}}
-
-\newcommand{\seal}[1]{\shareddom{seal}(#1)}
-\newcommand{\sealed}[1]{\shareddom{sealed}(#1)}
-
-\newcommand{\failed}{\mathrm{failed}}
-% DOMI: defining a macro named ``undefined'' breaks many latex packages.
-%       (this is just another way that LaTeX is broken as a programming language)
-% \newcommand{\undefined}{\mathrm{undefined}}
-\newcommand{\tundefined}{\mathrm{undefined}}
-\newcommand{\halted}{\mathrm{halted}}
-
-%%% Target domain
-\newcommand{\targetdom}[1]{\mathrm{#1}}
-\newcommand{\tRegName}{\targetdom{RegisterName}}
-
-%%% Programs and contexts
-\newcommand{\program}{\mathscr{P}}
-\newcommand{\context}{\mathscr{C}}
-
-\newcommand{\plug}[2]{#1[#2]}
-
-%%% Operational semantics
-\newcommand{\step}{\rightarrow}
-\newcommand{\nstep}[1][n]{\step_{#1}}
-\newcommand{\steps}{\step^*}
-\newcommand{\diverge}{{\Uparrow}}
-\newcommand{\term}[1][-]{{\Downarrow_{#1}}}
-
-%%% Variables
-\newcommand{\var}[1]{\mathit{#1}}
-\newcommand{\rn}{\var{rn}}
-\newcommand{\reg}{\var{reg}}
-\newcommand{\mem}{\var{mem}}
-\newcommand{\ms}{\var{ms}}
-\newcommand{\pc}{\var{pc}}
-\newcommand{\stk}{\var{stk}}
-\newcommand{\link}{\var{link}}
-\newcommand{\stkf}{\stk_{\var{frame}}}
-\newcommand{\ret}{\var{ret}}
-\newcommand{\data}{\var{data}}
-\newcommand{\code}{\var{code}}
-\newcommand{\priv}{\var{priv}}
-\newcommand{\opc}{\var{opc}}
-\newcommand{\odata}{\var{odata}}
-\newcommand{\vsc}{\var{sc}}
-\newcommand{\cb}{\vsc}
-\newcommand{\baddr}{\var{b}}
-\newcommand{\eaddr}{\var{e}}
-\newcommand{\aaddr}{\var{a}}
-\newcommand{\stdrng}{[\baddr,\eaddr]}
-
-%%% Constants
-\newcommand{\constant}[1]{\mathrm{#1}}
-\newcommand{\calllen}{\constant{call\_len}}
-\newcommand{\stkb}{\constant{stk\_base}}
-
-%%% Named registers
-\newcommand{\pcreg}{\mathrm{pc}}
-\newcommand{\rstk}{\mathrm{r}_\mathrm{stk}}
-\newcommand{\rO}{\mathrm{r}_\mathrm{ret}}
-\newcommand{\rret}{\rO}
-\newcommand{\rretc}{\mathrm{r}_\mathrm{ret code}}
-\newcommand{\rretd}{\mathrm{r}_\mathrm{ret data}}
-\newcommand{\rdata}{\mathrm{r}_\mathrm{data}}
-\newcommand{\rtmp}[1]{\mathrm{r}_\mathrm{t#1}}
-
-%%% locality
-%\newcommand{\plainlocality}[1]{\mathrm{#1}}
-%\newcommand{\glob}{\plainlocality{global}}
-%\newcommand{\local}{\plainlocality{local}}
-
-%%% linearity
-\newcommand{\plainlinearity}[1]{\mathrm{#1}}
-\newcommand{\linear}{\plainlinearity{linear}}
-\newcommand{\normal}{\plainlinearity{normal}}
-
-%%% Permissions
-\newcommand{\plainperm}[1]{\textsc{#1}}
-%\newcommand{\rwlx}{\plainperm{rwlx}}
-\newcommand{\rwx}{\plainperm{rwx}}
-\newcommand{\rx}{\plainperm{rx}}
-%\newcommand{\rwlxo}{\plainperm{rwlxo}}
-%\newcommand{\rwlo}{\plainperm{rwlo}}
-%\newcommand{\rwl}{\plainperm{rwl}}
-%\newcommand{\rwxo}{\plainperm{rwxo}}
-%\newcommand{\rwo}{\plainperm{rwo}}
-\newcommand{\rw}{\plainperm{rw}}
-%\newcommand{\rxo}{\plainperm{rxo}}
-\newcommand{\readonly}{\plainperm{r}}
-\newcommand{\ro}{\readonly}
-\newcommand{\noperm}{\plainperm{0}}
-%\newcommand{\nopermo}{\plainperm{0o}}
-%\newcommand{\enter}{\plainperm{e}}
-%\newcommand{\entero}{\plainperm{eo}}
-
-%%% Braces
-\newcommand{\comp}[1]{[#1]}
-
-%%% Functions
-\newcommand{\plainfun}[2]{
-  \ifthenelse{\equal{#2}{}}
-  {\mathit{#1}}
-  {\mathit{#1}(#2)}
-}
-
-%\newcommand{\isLoc}[1]{\plainfun{isLocal}{#1}}
-%\newcommand{\nonLoc}[1]{\plainfun{nonLocal}{#1}}
-%\newcommand{\opaquePerm}[1]{\plainfun{opaquePerm}{#1}}
-%\newcommand{\updPcPerm}[1]{\plainfun{updatePcPerm}{#1}}
-%\newcommand{\writeLocalAllowed}[1]{\plainfun{writeLocalAllowed}{#1}}
-\newcommand{\addressable}[1]{\plainfun{addressable}{#1}}
-\newcommand{\callCond}[1]{\plainfun{callCondition}{#1}}
-\newcommand{\decInstr}[1]{\plainfun{decodeInstruction}{#1}}
-\newcommand{\decPerm}[1]{\plainfun{decodePerm}{#1}}
-\newcommand{\encInstr}[1]{\plainfun{encodeInstruction}{#1}}
-\newcommand{\encPerm}[1]{\plainfun{encocePerm}{#1}}
-\newcommand{\encLin}[1]{\plainfun{encoceLin}{#1}}
-\newcommand{\encType}[1]{\plainfun{encodeType}{#1}}
-\newcommand{\exec}[1]{\plainfun{executable}{#1}}
-\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
-\newcommand{\isLinear}[1]{\plainfun{isLinear}{#1}}
-\newcommand{\linCons}[1]{\plainfun{linearityConstraint}{#1}}
-\newcommand{\noRetStkMs}[1]{\plainfun{noRetStk_{ms}}{#1}}
-\newcommand{\noRetStkReg}[1]{\plainfun{noRetStk_{reg}}{#1}}
-\newcommand{\nonExec}[1]{\plainfun{nonExecutable}{#1}}
-\newcommand{\nonLinear}[1]{\plainfun{nonLinear}{#1}}
-\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
-\newcommand{\range}[1]{\plainfun{range}{#1}}
-\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
-\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
-\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
-\newcommand{\sealAss}[1]{\plainfun{sealAssignment}{#1}}
-\newcommand{\updPcAddr}[1]{\plainfun{updatePc}{#1}}
-\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
-\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
-\newcommand{\xjumpResult}[3]{\plainfun{xjumpResult}{#1,#2,#3}}
+\input{preamble}
 

### 2018-01-25 fully-abstract-comp/ftlr-proof-cases.tex
half om jmpx case of ftlr.

\documentclass[a4paper]{article}
\usepackage[a4paper]{geometry}
\usepackage[utf8]{inputenc}

\input{preamble}

\title{FTLR proof cases}

\begin{document}
\maketitle
\begin{proof}
  Assume
\[
  \npair{(b,e)}\in \readCond{\lin,W}
\]
and show
\[
\npair{((\rx,\lin),\baddr,\eaddr,\aaddr),
       ((\rx,\lin),\baddr,\eaddr,\aaddr)} \in \lre(W)
\]
We write $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
Proceed by induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}
Let $n' \leq n$ be given and assume
\begin{enumerate}
\item $\npair[n']{[(c,c)]} \in \lrp(W)$
\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
\item $W \oplus W_R \oplus W_M$ is well-defined.
\end{enumerate}
Further let
\begin{itemize}
\item $\Phi_S = (\ms_S,\reg_S\updReg{\pcreg}{c},\stk,\ms_\stk)$
\item $\Phi_T = (\ms_T,\reg_T\updReg{\pcreg}{c})$
\end{itemize}
and show
\[
  \npair[n']{(\Phi_S,\Phi_T)} \in \lro(W \dots)
\]
Let $i \leq n'$ be given \footnote{ignoring the frame...} and show
\[
  \Phi_S \term[i] \Rightarrow \Phi_T \term
\]
and 
\[
  \Phi_T \term[i] \Rightarrow \Phi_S \term
\]
First show the former of the two, assume $\Phi_S\term[i]$. We now consider the first step of $\Phi_S \term[i]$\footnote{if $i=0$, then we have a contradiction with $\Phi_S\term[i]$.}. 
The possible instructions that could have been executed are:
\begin{enumerate}
\item $\txjmp{r_1}{r_2}$
\item $\dots$
\end{enumerate}
In the case $\txjmp{r_1}{r_2}$\footnote{$\pcreg$ is a corner case here. It will fail contradicting previous assumption of termination.} was executed, we know that $\Phi_S(r_1)$ and $\Phi_S(r_2)$ either points to two sealed capabilities or a return pointer pair (all other cases result in $\Phi_S \step \failed$ contradicting $\Phi_S\term[i]$). From assumption \ref{item:reg-ass} we know
\[
  \npair[n']{(\Phi_S(r_i), \Phi_T(r_i))} \in \lrv(W_{R,i}) \text{ for $i = 1,2$}
\]
where $W_{R,i}$ are partitions of the partitioning of $W_R$ we get from assumption \ref{item:reg-ass}. From the previous observation w.r.t. $\Phi_S(r_1)$ and $\Phi_S(r_2)$, we know that we are in the "sealed capability case" if $\lrv$. This in particular gives us that
\[
  \begin{array}{l}
    \forall W' \future W, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
    \quad \npair[n']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W_{R,1} \oplus W_o) \wedge \dots
  \end{array}
\]
and
\[
  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}
\]
where $\Phi_T(r_2) = \sealed(\sigma,\vsc_{T,2})$.
Now use the former with $W'=W_{R,1}$, $W_o = W_{R,2}$, and $n'' = i-1$ to get
\[
  \npair[n']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1} \oplus W_{R,2}) 
\]
Using $\lrexj$, we want to ``start from'' configuration pair $(\Phi_S,\Phi_T)$ and end up in $(\Phi_S',\Phi_T')$ where
\begin{itemize}
\item $\Phi_S' = \xjumpResult{\Phi_S(r_1)}{\Phi_S(r_2)}{\Phi_S}$
\item $\Phi_T' = \xjumpResult{\Phi_T(r_1)}{\Phi_T(r_2)}{\Phi_T}$
\end{itemize}
so we simply pick the same memory as in $\Phi_S,\Phi_T$ and the same register file but with $r_1$ and $r_2$ cleared if $\Phi(r_i)$ is linear, respectively. Using $W_M$ memory satisfaction follows from assumption \ref{item:mem-ass}. Using $W_R$ with all the ownership of regions owned by $W_{R,i}$, $i =1,2$ stripped, the register-file satisfaction follows from \ref{item:reg-ass}. Finally the $\npair[i-1]{[(\Phi_S(r_1),\Phi_T(r_1)),(\Phi_S(r_2),\Phi_T(r_2))]}\in\lrp(W_{R,1},W_{R,2})$ follows from the $\lrv$ assumption. Finally the consistency of the worlds we use follows easily.

All in all, this allows us to conclude
\[
  \npair[i-1]{\Phi_S',\Phi_T'} \in \lro(W_{R,1} \oplus \dots)
\]
From the assumption $\Phi_S\term[i]$, we get $\Phi_S\term[i-1]$ which we use with the above to conclude $\Phi_T' \term$. By the operational semantics of $\txjmp{r_1}{r_2}$, $\Phi_T \step \Phi_T'$ with which we conclude $\Phi_T \term$.

% Now show
% \[
%   \Phi_T \term[i] \Rightarrow \Phi_S \term
% \]

\end{proof}
\end{document}

### 2018-02-06 esop-2018/paper.tex
Add my acknowledgements to ESOP paper.
A couple of updates to meet the Springer guidelines.

@@
 
-\bibliographystyle{splncs}
+\bibliographystyle{MathPhySci}
 
@@
 \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
-\institute{Aarhus University - \email{\{lau,birkedal\}@cs.au.dk}
-\and imec-DistriNet, KU Leuven - \email{dominique.devriese@cs.kuleuven.be}}
+\institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
+\and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}
 
@@
 
-\section{A capability machine with local capabilities}
+\section{A Capability Machine with Local Capabilities}
 \label{sec:capab-mach-with}
@@
 
-\section{Stack and return pointer management using local capabilities}
+\section{Stack and Return Pointer Management Using Local Capabilities}
 \label{sec:stack-and-return-pointer} One of the contributions in this paper is a
@@
 
-\section{Logical relation}
+\section{Logical Relation}
 \label{sec:logical-relation}
@@
 
-\subsubsection{Future worlds}
+\subsubsection{Future Worlds}
 \label{subsec:future-worlds} 
@@
 
-\subsubsection{World satisfaction}
+\subsubsection{World Satisfaction}
 A memory satisfies a world, written $\memSat{\ms}{W}$, if it can be partitioned
@@
 
-\subsection{Logical relation}
+\subsection{Logical Relation}
 \afterpage{
@@
 
-\subsection{Safety of the capability machine}
+\subsection{Safety of the Capability Machine}
 With the logical relation defined, we can now state the fundamental theorem of
@@
 
-\subsection{Encapsulation of local state}
+\subsection{Encapsulation of Local State}
 \texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
@@
 
-\subsection{Well-bracketed control-flow} 
+\subsection{Well-Bracketed Control-Flow} 
 Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
@@
 
-\section{Related work}
+\section{Related Work}
 \label{sec:related-work}
@@
 from The Danish Council for Independent Research for the Natural Sciences (FNU).
+Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
 

### 2018-02-12 fully-abstract-comp/technical-report.tex
half om jmpx case of ftlr.
Fix typo.
Removed the world parameter from observation relation.
Added a new read-like condition for execution. The original readCondition is unchanged and it is assumed in conjuntion with the new readXCondition for executable capabilities.
New worlds (work in progress)

@@
     \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
-    &\Word & \defeq & \ints \uplus \SealableCaps\\
+    &\Word & \defeq & \ints \uplus \Caps\\
     \permbnf \in& \Perm & \defbnf & \dots \\
@@
 
-\paragraph{Termination computations}
+\paragraph{Terminating computations}
 We use the following notation to write that a configuration $\Phi$ successfully terminates:
@@
       \begin{array}[t]{rl}
-        \Phi' & \updReg{r_1}{w_1}\\
-              & \updReg{r_2}{0} \\
+        \Phi' & \updReg{r_2}{0} \\
               & \updReg{\pcreg}{\opc}\\
@@
 \[
-  \World = \RegionName \parfun \Region
+  \World = \Worldh \times \Worlds \times \Worldfs
+\]
+with $\Worlds$, $\Worldh$, and $\Region$ defined as follows
+\[
+  \Worldh = \RegionName \fun (\Regions + \Regionh)
+\]
+and
+\[
+  \Worlds = \RegionName \fun (\Regions \times \SealableCaps)
+\]
+and
+\[
+  \Worldfs = \RegionName \fun \Regions
 \]
@@
 \[
-  \Region = \left\{
-  \begin{array}{l}
-    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2}) \uplus \\
-    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2})\uplus \\
-    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \UPred{\MemSeg^2}) \uplus\\
+  \Regionh = 
+    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \URel{\MemSeg^2})
+\]
+and
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2})\uplus \\ 
     \{\revoked\}
@@
 \subsection{Memory satisfaction}
+{\color{green}
+Memory satisfaction for new worlds:
+\[
+  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
+      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
+      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
+      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
+      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W} \wedge \\
+      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W} \wedge \\
+      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W}
+    \end{array}
+  \right.
+\]
+
+\[
+  \memSatStack{\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      W_\var{stack} = \pi_2(W) \wedge \\
+      \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
+      \exists R_\ms : \dom(W_\var{stack}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
+      \quad \ms_T = \biguplus_{r \in \dom(W_\var{stack})} \pi_3(R_\ms(r)) \wedge \\
+      \quad \\
+      \quad \exists R_W : \dom(W_\var{stack}) \fun World \ldotp \\
+      \qquad \forall r \in \dom(W_\var{stack}), n' < n \ldotp \\
+      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(R))} \in W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r)) \\
+      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc
+
+    \end{array}
+  \right.
+\]
+
+\[
+  \memSatFStack{ms_\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      \dots
+    \end{array}
+  \right.
+\]
+
+\[
+  \memSatHeap{ms_\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      \dots
+    \end{array}
+  \right.
+\]
+
+}
+
 \[
@@
                        \endarray\right)}
-      \in \lro(W_{\var{new}})
+      \in \lro
     \end{array}
@@
       \qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\\
-      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro(W \oplus W_\lrr \oplus W_\lrm)
+      \qquad W \oplus W_\lrr \oplus W_\lrm \\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
     \end{array}
@@
       \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
-      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro(W \oplus W_\lrr \oplus W_\lrm)
+      \qquad W \oplus W_\lrr \oplus W_\lrm\\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
     \end{array}
@@
 \[
-  \lro(W) = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+  \lro = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
@@
                                              \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0,0\endarray\right]\right)
-                                    \endarray\right)} \in \lro(W_\var{new})
+                                    \endarray\right)} \in \lro
     \end{array} 
@@
         \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
-        \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
-        \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
+        \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
+                                   \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
+        \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
+                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
       \end{array}
@@
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
+\[
+  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\lin, W}, \ms \ldotp\\
+      \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
+      \quad W(r) \nequal \stareg[(\ms,\ms)]{\pur}
+    \end{array}
+  \right\}
+\]
 
@@
 \]
+
+% \[
+%   \staureg{v} = (v,*,=,=,H^{\mathrm{sta},\lrv}_{\stpair{\ms}{\ms}})
+% \]
+% \[
+%   H^{\mathrm{sta},\lrv}_{\stpair{\ms}{\ms}} = 
+%     \left\{ \npair{\stpair{\ms}{\ms}} \middle|
+%       \begin{array}{l}
+%         \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\src{\ms_S})} S(\aaddr) \wedge\\        
+%         \quad \forall \aaddr \in \dom(\ms) \ldotp \npair[n-1]{\src{\ms_S}(\aaddr),\ms_T(\aaddr)} \in \lrv(S(a))
+%       \end{array}
+% \right\}
+% \]
 
@@
       \begin{array}{l}
-        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}
+        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
+        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}
       \end{array}
@@
   \label{thm:full-abstraction}
-  For programs $\program_1$ and $\program_2$ that are reasonably compiled compiled, we have
+  For programs $\program_1$ and $\program_2$ that are reasonably compiled, we have
   \begin{gather*}

### 2018-02-14 fully-abstract-comp/technical-report.tex
Make the colors more readable.

@@
 \documentclass[a3paper]{article}
+\usepackage[svgnames]{xcolor}
 \usepackage[a3paper,margin=0.7in]{geometry}
@@
 \subsection{Future world}
-\subsubsection{\color{green} first thoughts on change to future world relation}
-{\color{green}
+\subsubsection{\color{DarkGreen} first thoughts on change to future world relation}
+{\color{DarkGreen}
 The public part of the future world relation stays the same:
@@
 \subsection{Memory satisfaction}
-{\color{green}
+{\color{DarkGreen}
 Memory satisfaction for new worlds:

### 2018-02-14 fully-abstract-comp/ftlr-proof-cases.tex
typo
.

@@
 \begin{document}
+
 \maketitle
+\section{Lemmas}
+In this section, I have listed lemmas that seem to be necessary for the FTLR proof.
+\begin{lemma}[$\lrv$ mono]
+  For all $n$, $w_1$, $w_2$, and $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$
+\end{lemma}
+
+\section{FTLR proof}
+By induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
 \begin{proof}
@@
 We write $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
-Proceed by induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}
+
 Let $n' \leq n$ be given and assume
@@
 \[
-  \npair[n']{(\Phi_S,\Phi_T)} \in \lro(W \dots)
+  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
 \]
@@
 \]
+
+\subsubsection*{Assuming source machine terminates}
 First show the former of the two, assume $\Phi_S\term[i]$. We now consider the first step of $\Phi_S \term[i]$\footnote{if $i=0$, then we have a contradiction with $\Phi_S\term[i]$.}. 
@@
 \item $\txjmp{r_1}{r_2}$
-\item $\dots$
+\item $\scall[o]{}{r_1}{r_2}$
 \end{enumerate}
+
+\paragraph{Case: $\scall[o]{}{r_1}{r_2}$}
+In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution.
+We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}.
+By assumption, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
+\[
+  \npair[n']{(\src{\reg_S}(r_i),\reg_T(r_i))} \in \lrv(W_{r,1}) \text{ for i = 1,2}
+\]
+We can further deduce that $\src{\reg_S}(r_i)$ for $i=1,2$ are sealed capabilities as the execution would otherwise fail. This gives us
+\[
+  \begin{array}{l}
+    \forall W' \future W_{R,1}, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
+    \quad \npair[n'']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W' \oplus W_o) \wedge \dots
+  \end{array}
+\]
+and
+\[
+  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}
+\]
+At this point we need to pick a suitable $W' \future W_{R,1}$.
+To this end, we need to know a bit about $W_{R,1}$.
+From the operational semantics, we know $\Phi_S(\rstk) = \stkptr{\perm,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$.
+From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $W_{R,\stk} \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
+Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}.
+As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $W_{R,1}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
+We pick $W'$ to be
+\[
+  \begin{array}{r l}
+    W' = W     & \updReg{r}{\revoked}\\
+               & \updReg{z}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
+               & \updReg{y}{\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa}}
+  \end{array}
+\]
+where $x$ and $y$ are fresh region names.
+To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t.
+$W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$.
+Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$).
+Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get
+\[
+  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W'_{R,2}
+\]
+which we use to get:
+\[
+  \npair[n'']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1}' \oplus W_{R,2}')
+\]
+In order to use this, we need to pick a register-file pair and suitable memories.
+We pick a $\Phi_S$ in $\lrexj$, so the $\Phi_S'$ in $\lrexj$ becomes the $\Phi_S'$ we assumed $\Phi_S$ steps to.
+Specifically
+\[
+  \Phi_S.\reg
+  \begin{array}[t]{l}
+    \update{r_1}{\linCons{\Phi_S(r_1)}}\\
+    \update{r_2}{\linCons{\Phi_S(r_2)}}\\
+    \update{\pcreg}{0} \\
+    \update{\rdata}{0} \\
+    \update{\rstk}{c_\stk}\\
+    \update{\rretc}{c_{\ret,c}}\\
+    \update{\rretd}{c_{\ret,d}}\\
+    \update{\rtmp{1}}{0}\\
+  \end{array}
+\]
+where
+\begin{itemize}
+\item $c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}$
+\item $c_{\ret,c} = \retptrc(\sigma',(\perm,\lin),\baddr,\eaddr,\aaddr+\calllen))$
+\item $c_{\ret,d} = \retptrd(\sigma',(\aaddr_\stk,\eaddr_\stk))$
+\end{itemize}
+and
+$\ms_S,((\dots,\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}) :: \stk),\ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]}$
+
+For the target side, we pick the register-file
+\[
+  \Phi_T.\reg
+  \begin{array}[t]{l}
+    \update{r_1}{\linCons{\Phi_T(r_1)}}\\
+    \update{r_2}{\linCons{\Phi_T(r_2)}}\\
+    \update{\pcreg}{0} \\
+    \update{\rdata}{0} \\
+    \update{\rstk}{c_\stk'}\\
+    \update{\rretc}{c_{\ret,c}'}\\
+    \update{\rretd}{c_{\ret,d}'}\\
+    \update{\rtmp{1}}{0}\\
+  \end{array}  
+\]
+where
+\begin{itemize}
+\item $c_\stk' = ((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1)$
+\item $c_{\ret,c}' = \sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)}$
+\item $c_{\ret,d}' = \sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}$
+\end{itemize}
+\lau{We do not actually check the linearity of the stack, but it should have been checked that the base address actually is the stack base address which would give us that it is linear.
+This is not an assumption we have formally stated at the moment!}
+and the same memory
+\[
+  \Phi_T.\ms_T\update{\aaddr_\stk}{42}
+\]
+We have to argue (1) that this register-file pair is in the register-file relation for a world $W_R'$ with the same ownership as $W_R$ and regions $z$ and $y$ are owned, and (2) the memory-pair satisfies the memory satisfaction relation for a world $W_M'$ (i.e.
+with the same ownership as $W_M$).
+
+Let's start with (2).
+Here we use that we took a standard region and replaced by two standard regions.
+This means that when we split up the memory, we can reuse the same splitting og memory and world (w.r.t memory) except for the memory segment that previously satsfied the standard region in question.
+This memory segment, we split in two, so each piece fit with one of the new regions.
+
+For (1), we need to consider the contents of all registers.
+For the registers that were not overwritten, their safety follows from the register-file relation assumption we start with along with monotonicity and non-expansiveness of the value relation.
+This is also the case for $r_i$, $i=1,2$ if the capability is linear and it is trivial otherwise (as it is $0$).
+$\rtmp{1}$ is also trivial as it is $0$.
+This leaves $\rstk$, $\rretc$, and $\rretd$.
+
+For $\rstk$, we make sure the $z$ region is owned in the world we use.
+Now we use that we had a standard region governing the stack to get that everything on there is in the value relation.
+We then use the monotonicity of the value relation to conclude that everything is in the value relation in this future world which satisfies the read and write condition.
+
+For $\rretc$ we pick $W'$ as our world.
+It should be the case that $H_\sigma \; \sigma' \; W''_o$ only contains one \emph{executable} pair, namely the one we consider.
+Let $W'' \future W$, $n''<n'$, and $(c_S',\vsc_T') \in H_\sigma \; \sigma' \; W''_o$ be given and show
+\[
+  \npair[n'']{(c_{\ret,c},c_S',\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)},\sealed{\sigma',\vsc_T})}
+\]
+Here we can assume $(c_S',\vsc_T') \in H_\sigma \; \sigma' \;W''_o$ is a data capability\footnote{if it is not, then the execution would fail.
+Further, the other case we would have to consider ($c_{\reg,c}$ as the data capability) would also fail.}.\footnote{{\color{red}This suggest that $H_\sigma$ should accept only stack capabilities and specifically non-empty ones.}}
+
+\paragraph{Case: $\protect\txjmp{r_1}{r_2}$} 
 In the case $\txjmp{r_1}{r_2}$\footnote{$\pcreg$ is a corner case here.
@@
 From the previous observation w.r.t.
-$\Phi_S(r_1)$ and $\Phi_S(r_2)$, we know that we are in the "sealed capability case" if $\lrv$.
+$\Phi_S(r_1)$ and $\Phi_S(r_2)$, we know that we are in the "sealed capability case" of $\lrv$.
 This in particular gives us that
@@
   \begin{array}{l}
-    \forall W' \future W, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
-    \quad \npair[n']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W_{R,1} \oplus W_o) \wedge \dots
+    \forall W' \future W_{r,1}, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
+    \quad \npair[n'']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W' \oplus W_o) \wedge \dots
   \end{array}
@@
 \]
-where $\Phi_T(r_2) = \sealed(\sigma,\vsc_{T,2})$.
+where $\Phi_T(r_2) = \sealed{\sigma,\vsc_{T,2}}$.
 Now use the former with $W'=W_{R,1}$, $W_o = W_{R,2}$, and $n'' = i-1$ to get
 \[
-  \npair[n']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1} \oplus W_{R,2}) 
+  \npair[i-1]{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1} \oplus W_{R,2}) 
 \]
@@
 \[
-  \npair[i-1]{\Phi_S',\Phi_T'} \in \lro(W_{R,1} \oplus \dots)
+  \npair[i-1]{\Phi_S',\Phi_T'} \in \lro
 \]
-From the assumption $\Phi_S\term[i]$, we get $\Phi_S\term[i-1]$ which we use with the above to conclude $\Phi_T' \term$.
+From the assumption $\Phi_S\term[i]$, we get $\Phi'_S\term[i-1]$ which we use with the above to conclude $\Phi_T' \term$.
 By the operational semantics of $\txjmp{r_1}{r_2}$, $\Phi_T \step \Phi_T'$ with which we conclude $\Phi_T \term$.
@@
 
+\subsubsection*{Assuming target machine terminates}
+\paragraph{Case: }
+In the case  is the first step of the execution,
+
 \end{proof}

### 2018-02-20 local-capability-machine/local-capability-machine.tex
typo

@@
 \newcommand\bnfdef{\mathrel{::=}}
-\title{Reasoning about a Capability Machine with Local Capabilities\\
- Provably Safe Stack and Return Pointer Management (without OS Support)\\
+\title{Reasoning About a Machine with Local Capabilities\\
+ Provably Safe Stack and Return Pointer Management\\
 Technical Appendix Including Proofs and Details}
+
+\author{Lau Skorstengaard\\Aarhus University\\\texttt{lask@cs.au.dk}
+        \and 
+        Dominique Devriese\\imec-DistriNet, KU Leuven\\\texttt{dominique.devriese@cs.kuleuven.be}
+        \and 
+        Lars Birkedal\\Aarhus University\\\texttt{birkedal@cs.au.dk}}
 
@@
   \[
-   \Rels = \{\phi_\pub \times \phi \in \powerset{\States^2}\times \powerset{\States^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
+   \Rels = \{(\phi_\pub, \phi) \in \powerset{\States^2}\times \powerset{\States^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
   \]

### 2018-02-23 fully-abstract-comp/technical-report.tex
.
Fix three technical issues in the world satisfaction that I just discussed with Lau.
Add a missing projection.
Added the rest of the memory satisfaction defintion.
removed the other value relation and xjmp. Made the format A4 as no definitions are too long for this.
updated the world sections. Hopefully, everything necessary for the new worlds is there (except a solution for the recursive domain equation).

@@
-\documentclass[a3paper]{article}
+\documentclass[a4paper]{article}
 \usepackage[svgnames]{xcolor}
-\usepackage[a3paper,margin=0.7in]{geometry}
+\usepackage[a4paper,margin=0.7in]{geometry}
 \usepackage[utf8]{inputenc}
@@
     \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
-    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
+    &\rbnf & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
     &\Reg & \defeq & \RegName \fun \Word\\
@@
 \]
+
+We now define
 
@@
 \lau{28-09-2017: For now, the jump to a return pointer does the stack end point check. We could possibly move this to the proposed semantic condition as well.}
-\begin{align*}
-  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
-                                   \begin{cases}
-                                     \arraycolsep=0pt
-                                     \begin{array}[t]{rl}
-                                       \Phi & \updReg{r_1}{w_1}\\
-                                            & \updReg{r_2}{w_2}\\
-                                            & \updReg{\pcreg}{c_\code}\\
-                                            & \updReg{\rdata}{c_\data}
-                                     \end{array} &
-                                     \begin{array}[t]{l}
-                                       \Phi(r_1) = \sealed{\sigma_1,c_\code} \tand \\
-                                       \Phi(r_2) = \sealed{\sigma_2,c_\data} \tand \\
-                                       \sigma_1 = \sigma_2 \tand \\
-                                       \nonExec{c_\data} \tand \\
-                                       w_1 = \linCons{\Phi(r_1)} \tand \\
-                                       w_2 = \linCons{\Phi(r_2)}
-                                     \end{array}
-                                     \\&\\
-                                     \sourcecolor\left.
-                                       \arraycolsep=0pt
-                                       \begin{array}[t]{rl}
-                                         \Phi' & \updReg{r_2}{0} \\
-                                               & \updReg{\pcreg}{\opc}\\
-                                               & \updReg{\rdata}{0}\\
-                                               & \updReg{\rstk}{c_\stk}\\
-                                               & \updReg{\rtmp{1}}{0}\\
-                                               & \updReg{\rtmp{2}}{0}
-                                       \end{array}\right.&
-                                     \sourcecolor\left.
-                                       \begin{array}[t]{l}
-                                         \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
-                                         \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
-                                         \sigma_1 = \sigma_2 \tand\\
-                                         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
-                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
-                                         \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
-                                         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
-                                         \eaddr_{\stk} + 1 = \aaddr_\stk \tand\\
-                                         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
-                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
-                                       \end{array}\right.
-                                     \\
-                                     \failed & \totherwise
-                                   \end{cases}
-\end{align*}
+% \begin{align*}
+%   \sem{\txjmp{r_1}{r_2}}(\Phi) = &
+%                                    \begin{cases}
+%                                      \arraycolsep=0pt
+%                                      \begin{array}[t]{rl}
+%                                        \Phi & \updReg{r_1}{w_1}\\
+%                                             & \updReg{r_2}{w_2}\\
+%                                             & \updReg{\pcreg}{c_\code}\\
+%                                             & \updReg{\rdata}{c_\data}
+%                                      \end{array} &
+%                                      \begin{array}[t]{l}
+%                                        \Phi(r_1) = \sealed{\sigma_1,c_\code} \tand \\
+%                                        \Phi(r_2) = \sealed{\sigma_2,c_\data} \tand \\
+%                                        \sigma_1 = \sigma_2 \tand \\
+%                                        \nonExec{c_\data} \tand \\
+%                                        w_1 = \linCons{\Phi(r_1)} \tand \\
+%                                        w_2 = \linCons{\Phi(r_2)}
+%                                      \end{array}
+%                                      \\&\\
+%                                      \sourcecolor\left.
+%                                        \arraycolsep=0pt
+%                                        \begin{array}[t]{rl}
+%                                          \Phi' & \updReg{r_2}{0} \\
+%                                                & \updReg{\pcreg}{\opc}\\
+%                                                & \updReg{\rdata}{0}\\
+%                                                & \updReg{\rstk}{c_\stk}\\
+%                                                & \updReg{\rtmp{1}}{0}\\
+%                                                & \updReg{\rtmp{2}}{0}
+%                                        \end{array}\right.&
+%                                      \sourcecolor\left.
+%                                        \begin{array}[t]{l}
+%                                          \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
+%                                          \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
+%                                          \sigma_1 = \sigma_2 \tand\\
+%                                          \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
+%                                          \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
+%                                          \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
+%                                          \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
+%                                          \eaddr_{\stk} + 1 = \aaddr_\stk \tand\\
+%                                          c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
+%                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
+%                                        \end{array}\right.
+%                                      \\
+%                                      \failed & \totherwise
+%                                    \end{cases}
+% \end{align*}
 
@@
                                                         &\updReg{\rstk}{c_\stk}\\
-                                                        &\updReg{\rretc}{\retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))}\\
+                                                        &\left[
+\arraycolsep=0pt\array{l}
+\reg.\rretc \mapsto \\
+\hfill \retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))
+\endarray
+\right]\\
                                                         &\updReg{\rretd}{\retptrd(\sigma_?,(\aaddr_\stk,\eaddr_\stk))}\\
@@
 \[
-  \Worldh = \RegionName \fun (\Regions + \Regionh)
+  \Worldh = \RegionName \parfun (\Regions + \Regionh)
 \]
@@
 \[
-  \Worlds = \RegionName \fun (\Regions \times \SealableCaps)
+  \Worlds = \RegionName \parfun (\Regions \times \SealableCaps)
 \]
@@
 \[
-  \Worldfs = \RegionName \fun \Regions
+  \Worldfs = \RegionName \parfun \Regions
 \]
@@
 Notice that no region allows for both linear and non-linear capabilities to address it.
-\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
+
 \[
@@
 
-Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions.
+We introduce a bit of notation for projecting out each part of the world:
+\[
+  \begin{array}{l}
+    \pwheap = \pi_1(W)\\
+    \pwpriv = \pi_2(W)\\
+    \pwfree = \pi_3(W)
+  \end{array}
+\]
+as well as projections for the regions:
+\[
+  \begin{array}{l}
+  \prv{(v,s,\phi_\pub,\phi,H)} = v \\
+  \prv{\revoked} = \revoked     
+  \end{array}
+\]
+
+We define erasure for worlds as follows:
+\[
+\erase{(W_{\mathrm{heap}},W_{\mathrm{priv}},W_{\mathrm{free}})}{S} =
+  \left( \erase{W_{\mathrm{heap}}}{S},\erase{W_{\mathrm{priv}}}{S},\erase{W_{\mathrm{free}}}{S} \right)
+\]
+where erasure for each part of a world is defined as follows:
+\[
+\erase{W_{\mathrm{heap}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{W_{\mathrm{heap}}(r)} \in S\\
+      \bot
+    \end{array}
+  \right.
+\]
+\[
+\erase{W_{\mathrm{priv}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{\pregion{W_{\mathrm{priv}}(r)}} \in S \\
+      \bot
+    \end{array}
+  \right.
+\]
+\[
+\erase{W_{\mathrm{free}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{W_{\mathrm{free}}(r)} \in S\\
+      \bot
+    \end{array}
+  \right.
+\]
+
+The $\activeReg{}$ function takes a world and filters away all the revoked regions, so
+\[
+  \activeReg{W} = \erase{W}{\spatial,\spatialo,\pure}
+\]
+
+Disjoint union of worlds.
+Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions.
+Two worlds can be joined together if all of their three parts agree on the region names and each of their regions can be joined together.
 \[
   W_1 \oplus W_2 = W
-  \begin{array}[t]{l}
-    \text{if $\dom(W) = \dom(W_1) = \dom(W_2) \tand$} \\
-    \forall r \in \dom(W) \ldotp W(r) = W_1(r) \oplus W_2(r)
+  \text{ iff }
+  \begin{array}[t]{l}
+    \text{$\dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand$} \\
+    \text{$\dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand$} \\
+    \text{$\dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand$} \\
+    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
+    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
+    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
   \end{array}
@@
 \begin{mathpar}
-  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \pubft W(r) }
+  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \pubft W.i(r) }
             { W' \future W }
@@
 \begin{mathpar}
-  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \privft W(r) \\  \exists r \in \dom(W') \setminus \dom(m(W))\ldotp W'(r) \text{ is non-empty and stack region}}
+  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \privft W.i(r) \\  \exists r \in \dom(W'.\mathrm{priv}) \setminus \dom(m_{\mathrm{priv}}(\pwpriv))\ldotp \pwpriv[W'](r) \text{ is non-empty}}
             { W' \future W }
@@
 
-\begin{definition}
-  We say a region $(k,s,\phi_\pub,\phi,H)$ is a stack region on a machine with starting stack domain $[\stkb,\eaddr]$ iff
-\[
-  \begin{array}{l}
-    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) \subseteq [\stkb,\eaddr]  \wedge \dom(\ms_T) \subseteq [\stkb,\eaddr]
-  \end{array}
-\]
-\end{definition}
 }
 
-For both $\privft$ and $\pubft$ we have
-\begin{mathpar}
-  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
-            { W' \future W }
-\end{mathpar}
+% For both $\privft$ and $\pubft$ we have
+% \begin{mathpar}
+%   \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
+%             { W' \future W }
+% \end{mathpar}
 %
@@
 \subsection{Memory satisfaction}
+\lau{14-02-2018: No matter what the memory satisfaction ends up looking like, remember to handle the revoked case.}
 {\color{DarkGreen}
@@
       \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
+      W = W_{\var{stack}} \uplus W_{\var{free\_stack}} \uplus W_{\var{heap}} \wedge\\
       \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
       \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
-      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W} \wedge \\
-      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W} \wedge \\
-      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W}
+      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
+      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
+      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W_{\var{heap}}}
     \end{array}
@@
       \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
-      \exists R_\ms : \dom(W_\var{stack}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
-      \quad \ms_T = \biguplus_{r \in \dom(W_\var{stack})} \pi_3(R_\ms(r)) \wedge \\
-      \quad \\
-      \quad \exists R_W : \dom(W_\var{stack}) \fun World \ldotp \\
-      \qquad \forall r \in \dom(W_\var{stack}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(R))} \in W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r)) \\
-      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc
-
+      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World \ldotp \\
+      \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
+      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r)) \\
+      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
+      \qquad \quad \exists i \ldotp \opc_i = W_\var{stack}(r).\opc \wedge \ms_i = \pi_1 (R_\ms(r))
     \end{array}
@@
     \begin{array}{l}
-      \dots
+      W_\var{stack} = \pi_3(W) \wedge \\
+      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms_\stk = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
+      \qquad W = \oplus_{r \in \activeReg{W_\var{stack}}} R_W(r) \wedge\\
+      \qquad \forall r \in \activeReg{W_\var{stack}}, n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r))
     \end{array}
@@
 \[
-  \memSatHeap{ms_\stk,\ms_T}{W} \text{ iff } 
+  \memSatHeap{ms,\ms_T}{W} \text{ iff } 
   \left\{
     \begin{array}{l}
-      \dots
+      W_\var{heap} = \pi_1(W) \wedge \\
+      \exists R_\ms : \dom(\activeReg{W_\var{heap}}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{heap}})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms = \biguplus_{r \in \dom(\activeReg{W_\var{heap}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{W_\var{heap}}) \fun \World\ldotp\\
+      \qquad W = \oplus_{r \in \activeReg{W_\var{heap}}} R_W(r) \wedge\\
+      \qquad \forall r \in \activeReg{W_\var{heap}}, n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{heap}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r))
     \end{array}
@@
 \]
-
 }
 
-\[
-  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
-  \left\{
-    \begin{array}{l}
-      \exists R_\ms : \dom(W) \fun \MemSeg^2 \ldotp \\
-      \quad \ms_T = \biguplus_{r \in \dom(W)} \pi_2(R_\ms(r)) \wedge \\
-      \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
-      \quad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \uplus \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
-      \quad \exists R_w : \dom(W) \fun \World \ldotp\\
-      \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
-      \qquad \forall r \in \dom(W) , n' < n \ldotp \\
-      \quad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
-    \end{array}
-  \right.
-\]
-In the definition of memory satisfaction, $R_\ms$ creates a partitioning of the two related memories.
+% \[
+%   \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
+%   \left\{
+%     \begin{array}{l}
+%       \exists R_\ms : \dom(W) \fun \MemSeg^2 \ldotp \\
+%       \quad \ms_T = \biguplus_{r \in \dom(W)} \pi_2(R_\ms(r)) \wedge \\
+%       \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
+%       \quad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \uplus \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
+%       \quad \exists R_w : \dom(W) \fun \World \ldotp\\
+%       \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
+%       \qquad \forall r \in \dom(W) , n' < n \ldotp \\
+%       \quad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
+%     \end{array}
+%   \right.
+% \]
+% In the definition of memory satisfaction, $R_\ms$ creates a partitioning of the two related memories.
 As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language.
 We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints imposed by the world are satisfied.
-\lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
-\dominique{4-12-2017: Regarding the question whether $\opc$ should be mentioned in the memory satisfaction relation.
-  The relevant intuitive aspect of the proof here, I think, is that on a return, the source semantics restores opc from the copy kept in the stack, while the target semantics restores it from the copy in the code part of the return pointer.
-  We will need to prove (somewhere) that these two correspond, which means we will need to track of the invariant that guarantees this somewhere (perhaps here?).
-  I suspect the relevant invariants are (1) that any executable pointer signed with $\sealAss{a}$ must point to $a + \calllen$ and (2) that anything related to a $\rretc$ must be executable and sealed.
-  Perhaps this invariant is easiest to track in (1) the general requirement on the $H_\sigma$ and (2) in the value relation.
-  This suggests nothing should be required about opc here?
-  To be confirmed in the proof of the fundamental theorem.
-}
+% \lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
+% \dominique{4-12-2017: Regarding the question whether $\opc$ should be mentioned in the memory satisfaction relation.
+%   The relevant intuitive aspect of the proof here, I think, is that on a return, the source semantics restores opc from the copy kept in the stack, while the target semantics restores it from the copy in the code part of the return pointer.
+%   We will need to prove (somewhere) that these two correspond, which means we will need to track of the invariant that guarantees this somewhere (perhaps here?).
+%   I suspect the relevant invariants are (1) that any executable pointer signed with $\sealAss{a}$ must point to $a + \calllen$ and (2) that anything related to a $\rretc$ must be executable and sealed.
+%   Perhaps this invariant is easiest to track in (1) the general requirement on the $H_\sigma$ and (2) in the value relation.
+%   This suggests nothing should be required about opc here?
+%   To be confirmed in the proof of the fundamental theorem.
+% }
 
@@
 \lau{TODO: Rewrite the other definitions to pass around $H_\sigma$ (if this is how we want to handle seals.)}
-\[
-  \lrv(H_\sigma,W) =
-  \begin{array}[t]{l}
-    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
-%
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\
-                         \sealed{\sigma,((\perm,\lin),\baddr,\eaddr,\aaddr + ?)}\endarray\right)} \;\middle| \;
-    \begin{array}{l}
-      (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
-      \npair{([\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})]} \in \lrp(W)
-    \end{array}
-    \right\}\cup\\
-%
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,(\baddr,\eaddr))},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)} \;\middle| \;
-    \begin{array}{l}
-      \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
-      \forall n' \leq n \ldotp \forall W' \privft W , W_C,W_\lrm,W_\lrr\ldotp \\
-      \quad \npair[n']{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
-      \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\
-                                                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
-      \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
-      \quad \opc = ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C) \wedge \\
-      \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
-      \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
-      \quad \reg_T(r_\stk) = ((\rw,\linear),\stkb,\baddr-1,\_) \wedge \\
-      \quad W'(r_\var{free}) = \stdreg{[\stkb,\baddr-1]}{\spa} \wedge \\
-      \quad W'(r_\var{priv}) = \stareg[\src{\ms_f},\ms_f]{\spao} \wedge \\
-      \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
-      \quad W_\var{new} = (W' \oplus W_\lrr \oplus W_\lrm \oplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
-      \qquad \Rightarrow \\
-      \qquad \quad \npair[n']{\left(\arraycolsep=0pt\array{l}
-                                     \left(\arraycolsep=0pt\array{l}\ms_S,\reg_S \left[
-      \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0,0\endarray \right]\\
-                                     ,\stk,\ms_\stk \uplus \ms_f
-                                     \endarray \right),\\
-                                             \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0,0\endarray\right]\right)
-                                    \endarray\right)} \in \lro
-    \end{array} 
-    \right\}\cup \\
-%
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
-    \begin{array}{l}
-      \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
-      \forall W' \pubft W, n' \leq n \ldotp \\
-      \quad \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
-      \qquad \npair[n']{(\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T')} \in \lre(W') \wedge \\
-      \qquad \npair[n']{(\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T)} \in \lre(W')
-    \end{array}
-    \right\}\cup\\
-%
-    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+% \[
+%   \lrv(H_\sigma,W) =
+%   \begin{array}[t]{l}
+%     \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+% %
+%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\
+%                          \sealed{\sigma,((\perm,\lin),\baddr,\eaddr,\aaddr + ?)}\endarray\right)} \;\middle| \;
+%     \begin{array}{l}
+%       (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
+%       \npair{([\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})]} \in \lrp(W)
+%     \end{array}
+%     \right\}\cup\\
+% %
+%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,(\baddr,\eaddr))},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)} \;\middle| \;
+%     \begin{array}{l}
+%       \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
+%       \forall n' \leq n \ldotp \forall W' \privft W , W_C,W_\lrm,W_\lrr\ldotp \\
+%       \quad \npair[n']{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
+%       \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\
+%                                                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
+%       \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+%       \quad \opc = ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C) \wedge \\
+%       \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
+%       \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
+%       \quad \reg_T(r_\stk) = ((\rw,\linear),\stkb,\baddr-1,\_) \wedge \\
+%       \quad W'(r_\var{free}) = \stdreg{[\stkb,\baddr-1]}{\spa} \wedge \\
+%       \quad W'(r_\var{priv}) = \stareg[\src{\ms_f},\ms_f]{\spao} \wedge \\
+%       \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
+%       \quad W_\var{new} = (W' \oplus W_\lrr \oplus W_\lrm \oplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
+%       \qquad \Rightarrow \\
+%       \qquad \quad \npair[n']{\left(\arraycolsep=0pt\array{l}
+%                                      \left(\arraycolsep=0pt\array{l}\ms_S,\reg_S \left[
+%       \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0,0\endarray \right]\\
+%                                      ,\stk,\ms_\stk \uplus \ms_f
+%                                      \endarray \right),\\
+%                                              \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0,0\endarray\right]\right)
+%                                     \endarray\right)} \in \lro
+%     \end{array} 
+%     \right\}\cup \\
+% %
+%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
+%     \begin{array}{l}
+%       \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
+%       \forall W' \pubft W, n' \leq n \ldotp \\
+%       \quad \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
+%       \qquad \npair[n']{(\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T')} \in \lre(W') \wedge \\
+%       \qquad \npair[n']{(\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T)} \in \lre(W')
+%     \end{array}
+%     \right\}\cup\\
+% %
+%     % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+%     % \right\} \cup \\
+% %
+%     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+%     \begin{array}{r l l}
+%       \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
+%       \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
+%     \end{array}
 %     \right\} \cup \\
-%
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
-    \begin{array}{r l l}
-      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
-      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
-    \end{array}
-    \right\} \cup \\
-%
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
-    \begin{array}{r l l }
-      \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
-      \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
-      \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
-      \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
-    \end{array}
-    \right\}
-  \end{array}
-\]
+% %
+%     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+%     \begin{array}{r l l }
+%       \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
+%       \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
+%       \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
+%       \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
+%     \end{array}
+%     \right\}
+%   \end{array}
+% \]
+
 where $H_\sigma$ is a function from seals to the set of pairs of sealable capabilities that can be sealed with $\sigma$.

### 2018-02-26 fully-abstract-comp/technical-report.tex
Comment out old expression relation and execute condition and FTLR and comments too, now that the old value relation has been commented out.

@@
 
-\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
 It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
@@
 \lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
-\[
-  \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
-    \begin{array}{l}
-      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \forall W_\lrr, W_\lrm \ldotp \\
-      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge\\
-      \qquad W_{\var{new}} = W \oplus W_\lrr \oplus W_\lrm \\
-      \qquad\Rightarrow\\
-      \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata}{v_{c,S},v_{d,S}},\stk, \ms_\stk)},\\
-                                          (\ms_T,\reg_T\updReg{\pcreg,\rdata}{v_{c,T},v_{d,T}})
-                       \endarray\right)}
-      \in \lro
-    \end{array}
-    \right\}
-\]
-\lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out.
+% \[
+%   \lre(W) = \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+%     \begin{array}{l}
+%       \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+%       \quad \forall W_\lrr, W_\lrm \ldotp \\
+%       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
+%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+%       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge\\
+%       \qquad W_{\var{new}} = W \oplus W_\lrr \oplus W_\lrm \\
+%       \qquad\Rightarrow\\
+%       \qquad\npair[n']{\left(\array{l}\src{(\ms_S,\reg_S\updReg{\pcreg,\rdata}{v_{c,S},v_{d,S}},\stk, \ms_\stk)},\\
+%                                           (\ms_T,\reg_T\updReg{\pcreg,\rdata}{v_{c,T},v_{d,T}})
+%                        \endarray\right)}
+%       \in \lro
+%     \end{array}
+%     \right\}
+% \]
+% \lau{24-10-2017: Dominique, I do not see how the splitting of worlds is going to work out.
 I can see how it will enforce linearity, but if you give a "spatial" region (the kind of region that is only allowed to go to one part of a split) to the $\lrr$-relation, then the memory satisfaction does not have it, and it cannot find a memory segment for this part.
@@
 Maybe the memory satisfaction should have all of the parts, so it can represent them (but also limit what linear capabilities can be used).}
-\dominique{18-1-2018: Lau, I suppose the above has been answered verbally already during one of the previous skype calls, but for reference, this is what the spatial vs spatial\_owned distinction is intended to solve: the world will contain spatial regions for such memory segments, but not spatial\_owned ones, so the memory satisfaction will require the memory to be present but not allow it to be addressed.}
-
-\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+% \dominique{18-1-2018: Lau, I suppose the above has been answered verbally already during one of the previous skype calls, but for reference, this is what the spatial vs spatial\_owned distinction is intended to solve: the world will contain spatial regions for such memory segments, but not spatial\_owned ones, so the memory satisfaction will require the memory to be present but not allow it to be addressed.}
+
+% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
 It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
@@
 
-\[
-  \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
-    \begin{array}{l}
-      \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
-      \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
-      \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}]} \in \lrp(W' \oplus W'') \Rightarrow \\
-      \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
-    \end{array}
-    \right\}
-\]
-
-\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+% \[
+%   \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
+%     \begin{array}{l}
+%       \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
+%       \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
+%       \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}]} \in \lrp(W' \oplus W'') \Rightarrow \\
+%       \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
+%     \end{array}
+%     \right\}
+% \]
+
+% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
 It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
@@
 
-\begin{theorem}[FTLR]
-  \label{thm:ftlr}
-  \[
-    \begin{array}{l}
-      \forall n,W,W_D,W_C,\perm_C,\perm_D,\lin_C,\lin_D,\baddr_C,\baddr_D,\eaddr_C,\eaddr_D,\aaddr_C,\aaddr_D \ldotp \\
-      \quad \left(
-      \begin{array}{l}
-        W = W_D \oplus W_C \wedge \\
-        \perm_D \in \readAllowed{} \Rightarrow \npair{(\baddr_D,\eaddr_D)} \in \readCond{\lin_D,W_D} \wedge\\
-        \perm_D \in \writeAllowed{} \Rightarrow\npair{(\baddr_D,\eaddr_D)} \in \writeCond{\lin_D,W_D} \wedge\\
-        \perm_D \notin \{\rwx,\rx\} \wedge \\
-        \perm_C = \rx \wedge\\
-        \npair{(\baddr_C,\eaddr_C)} \in \readCond{\lin_C,W_C}
-      \end{array}
-      \right)\\
-      \quad\Rightarrow 
-      \npair{\left(\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
-                      ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray \right)} \in \lre(W)
-
-    \end{array}
-  \]
-\end{theorem}
-
-\dominique{22-1-2018: the below is an alternative proposal to the above definition.
+% \begin{theorem}[FTLR]
+%   \label{thm:ftlr}
+%   \[
+%     \begin{array}{l}
+%       \forall n,W,W_D,W_C,\perm_C,\perm_D,\lin_C,\lin_D,\baddr_C,\baddr_D,\eaddr_C,\eaddr_D,\aaddr_C,\aaddr_D \ldotp \\
+%       \quad \left(
+%       \begin{array}{l}
+%         W = W_D \oplus W_C \wedge \\
+%         \perm_D \in \readAllowed{} \Rightarrow \npair{(\baddr_D,\eaddr_D)} \in \readCond{\lin_D,W_D} \wedge\\
+%         \perm_D \in \writeAllowed{} \Rightarrow\npair{(\baddr_D,\eaddr_D)} \in \writeCond{\lin_D,W_D} \wedge\\
+%         \perm_D \notin \{\rwx,\rx\} \wedge \\
+%         \perm_C = \rx \wedge\\
+%         \npair{(\baddr_C,\eaddr_C)} \in \readCond{\lin_C,W_C}
+%       \end{array}
+%       \right)\\
+%       \quad\Rightarrow 
+%       \npair{\left(\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
+%                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray \right)} \in \lre(W)
+
+%     \end{array}
+%   \]
+% \end{theorem}
+
+% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
 It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
-Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps.
+% Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps.
 It expresses that any reasonable executable capability is safe to use as the pc capability.
-\dominique{I'm following the example of the above FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
+\dominique{I'm following the example of the previous FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
 \begin{theorem}[FTLR]

### 2018-02-26 fully-abstract-comp/technical-report.tex
Fix some typos.
Minor tweaks, remove outdated comment.
Add comment.
Add date to comment.
Fix sentence.

@@
     \begin{array}{l}
-      W_\var{stack} = \pi_2(W) \wedge \\
+      W_\var{stack} = \pwpriv \wedge \\
       \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
@@
     \begin{array}{l}
-      W_\var{stack} = \pi_3(W) \wedge \\
+      W_\var{stack} = \pwfree \wedge \\
       \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \MemSeg \wedge \\
@@
       \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
-      \qquad W = \oplus_{r \in \activeReg{W_\var{stack}}} R_W(r) \wedge\\
-      \qquad \forall r \in \activeReg{W_\var{stack}}, n' < n \ldotp \\
+      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
       \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r))
@@
     \begin{array}{l}
-      W_\var{heap} = \pi_1(W) \wedge \\
+      W_\var{heap} = \pwheap \wedge \\
       \exists R_\ms : \dom(\activeReg{W_\var{heap}}) \fun \MemSeg \times \MemSeg \wedge \\
@@
 \subsection{Relation}
-\lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
+% \lau{19-10-2017: Rest of source configuration needs to come from somewhere in $\lre$.}
 % \[
@@
   \]
+  \dominique{26-2-2018: $\readCond{\lin,W}$ and $\xReadCond{\lin,W}$ seem contradictory. I suspect the former needs to go?}
 \end{theorem}
@@
 \section{Back translation}
-The back translation is an embedding of source language into .
+The back translation is an embedding of the target language into the source language.
 

### 2018-03-06 fully-abstract-comp/ftlr-proof-cases.tex
Add placeholders for missing cases in FTLR proof.
Prove trivial case.
updated part of proof to use new worlds

@@
 \usepackage[utf8]{inputenc}
+
+\usepackage[svgnames]{xcolor}
 
@@
   For all $n$, $w_1$, $w_2$, and $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$
+\end{lemma}
+
+\begin{lemma}[Steps-to termination]
+  For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
+\[
+  \Phi_T \steps[j] \Phi_T' \text{ and } \npair{(\Phi_S,\Phi_T')} \in \lro,
+\]
+then
+\[
+  \npair{(\Phi_S,\Phi_T)} \in \lro
+\]
 \end{lemma}
@@
 \]
-At this point we need to pick a suitable $W' \future W_{R,1}$.
+At this point, we need to pick a suitable $W' \future W_{R,1}$.
 To this end, we need to know a bit about $W_{R,1}$.
@@
   \begin{array}{r l}
-    W' = W     & \updReg{r}{\revoked}\\
-               & \updReg{z}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
-               & \updReg{y}{\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa}}
+    W' = W     & \update{\mathrm{free}.r}{\revoked}\\
+               & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
+               & \update{\mathrm{priv}.y}{\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa}}
   \end{array}
 \]
-where $x$ and $y$ are fresh region names.
+where $x$ and $y$ are fresh region names in the respective part of the world.
 To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t.
@@
 \[
-  \npair[n'']{(c_{\ret,c},c_S',\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)},\sealed{\sigma',\vsc_T})}
+  \npair[n'']{(c_{\ret,c},c_S',\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)},\sealed{\sigma',\vsc_T})} \in \lrexj(W'')
 \]
-Here we can assume $(c_S',\vsc_T') \in H_\sigma \; \sigma' \;W''_o$ is a data capability\footnote{if it is not, then the execution would fail.
+Here we can assume $(c_S',\vsc_T') \in H_\sigma \; \sigma' \;W''_o$ is a data capability pair\footnote{if it is not, then the execution would fail.
 Further, the other case we would have to consider ($c_{\reg,c}$ as the data capability) would also fail.}.\footnote{{\color{red}This suggest that $H_\sigma$ should accept only stack capabilities and specifically non-empty ones.}}
+
+{\color{DarkRed}
+  A this point, we would like to use the induction hypothesis with the assumptions we get from $\lrexj$. That is, we use the induction hypothesis to (hopefully) get
+
+\[
+  \npair[n'']{((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots),((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen)} \in \lre(W'')
+\]
+(notice that the target has ``caught up'' here.) The assumptions we get from $\lrexj(W'')$ are not synced, so this would indicate that we need the target side to catch up.
+}
 

### 2018-03-07 fully-abstract-comp/technical-report.tex
small fix to call implementation
Added a bit more to memory satisfaction. Added comment to remember to update the parts of the LR that 'open' the world
typo.
Update comment.
Updated lro to be assymetric
Update read and write conditions for new worlds.

@@
 \end{align*}
+\dominique{Note: The above does not allow executing code that is on the stack in the source machine, while this is in principle allowed in the target.
+This is probably fine if we make sure a stack pointer can never be executable.}
 
@@
     \text{// push 42 on the stack (so it is non-empty).}\\
-    \tstore{\rstk}{42}\\
+    \tmove{\rtmp{1}}{42}\\
+    \tstore{\rstk}{\rtmp{1}}\\
     \tcca{\rstk}{(-1)}\\
@@
       \quad \exists R_W : \dom(\activeReg{W_\var{heap}}) \fun \World\ldotp\\
-      \qquad W = \oplus_{r \in \activeReg{W_\var{heap}}} R_W(r) \wedge\\
+      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{heap}})} R_W(r) \wedge\\
       \qquad \forall r \in \activeReg{W_\var{heap}}, n' < n \ldotp \\
@@
       \begin{array}{r l l}
-        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W} \wedge \\
-        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
+        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackWriteCond{W}
       \end{array}
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, W} \ldotp \\
+      \exists r \in \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
+      \qquad W(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
+\[
+  \stackReadCond{W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\linear, \pwfree} \ldotp \\
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, W} \ldotp \\
+      \exists r \in \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
+      \qquad W(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
+      \qquad W(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
+\[
+  \stackWriteCond{W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\linear, \pwfree} \ldotp \\
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
@@
   \]
-  \dominique{26-2-2018: $\readCond{\lin,W}$ and $\xReadCond{\lin,W}$ seem contradictory. I suspect the former needs to go?}
+  \dominique{6-3-2018: $\readCond{\lin,W}$ and $\xReadCond{\lin,W}$ seem contradictory.
+    I suspect the latter needs to go? I suspect we don't need to assume the context has no way to modify its own code.}
 \end{theorem}

### 2018-03-07 fully-abstract-comp/ftlr-proof-cases.tex
Some work on a lemma about the read and write condition. Add some remarks about problems with spurious later modalities.

@@
 \begin{lemma}[$\lrv$ mono]
-  For all $n$, $w_1$, $w_2$, and $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$
+  For all $n$, $w_1$, $w_2$, and $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$.
 \end{lemma}
@@
 \end{lemma}
+
+\begin{lemma}
+  If
+  \begin{itemize}
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
+  \item $\npair{(b,e)}\in\readCond{l,W}$
+  \item $a \in [b,e]$
+  \item $W \oplus W_M$ is defined
+  \item $n' < n$
+  \end{itemize}
+  Then $\npair[n']{(\ms_S(a),\ms_T(a))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+  Additionally, if
+  \begin{itemize}
+  \item $\npair{(b,e)}\in\writeCond{l,W}$
+  \end{itemize}
+  Then $\memSat{(\ms_S[a\mapsto 0],\stk,\ms_\stk,\ms_T[a\mapsto 0])}{W_M'}$.
+\end{lemma}
+\begin{proof}
+  From $\npair{(b,e)}\in\readCond{l,W}$, we get an $r \in \addressable{\lin,
+    \pwheap}$ and $[\baddr',\eaddr'] \supseteq [\baddr,\eaddr]$ such that $\pwheap(r)
+  \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}$.
+
+  Since $W \oplus W_M$ is defined, we have that $r \in \dom(\pwheap) =
+  \dom(\pwheap[W_M])$ and $\pwheap(r) \oplus \pwheap[W_M](r)$ is defined.
+
+  From $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$, we get that
+  $\stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m)$,
+  $\ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m$,
+  $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
+  $\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}$ such that
+  \begin{itemize}
+  \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus
+    \ms_\var{T,heap}$
+  \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
+  \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
+  \item $\memSatHeap{\ms_S,\ms_\var{T,heap}}{W_\var{heap}}$.
+  \end{itemize}
+
+  From $\memSatHeap{\ms_S,\ms_\var{T,heap}}{W_\var{heap}}$, we get an
+  $R_\ms : \dom(\activeReg{\pwheap[W_\var{heap}]}) \fun \MemSeg \times \MemSeg$,
+  $\ms_\var{T,heap} = \biguplus_{r \in \dom(\activeReg{\pwheap[W_\var{heap}]})} \pi_2(R_\ms(r))$,
+  $\ms_S = \biguplus_{r \in \dom(\activeReg{\pwheap[W_\var{heap}]})} \pi_1(R_\ms(r))$,
+  $\exists R_W : \dom(\activeReg{\pwheap[W_\var{heap}]}) \fun \World\ldotp$
+  $W_\var{heap} = \oplus_{r \in \dom(\activeReg{\pwheap[W_\var{heap}]})} R_W(r)$,
+  $\forall r \in \activeReg{\pwheap[W_\var{heap}]}, n' < n$, we have that
+  $\npair[n']{R_\ms(r)} \in  \pwheap[W_\var{heap}](r).H \; \pwheap[W_\var{heap}](r).s \; \xi^{-1}(R_W(r))$.
+
+  We have that $r \in \addressable{\lin, \pwheap} \subseteq
+  \activeReg{\pwheap[W_\var{heap}]}$, so $\npair[n']{R_\ms(r)} \in
+  \pwheap[W_\var{heap}](r).H \; \pwheap[W_\var{heap}](r).s \; \xi^{-1}(R_W(r))$.
+  Because $\pwheap(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
+  also $\pwheap[W_\var{heap}](r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}$ this
+  means that also $\npair[n']{R_\ms(r)} \in
+  H^{\mathrm{std}}_{[\baddr',\eaddr']}\; \xi^{-1}(R_W(r))$.
+
+  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = [\baddr',\eaddr']$ and we get a $S : [\baddr',\eaddr'] \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in [\baddr',\eaddr']} S(\aaddr)$ and $\forall \aaddr \in [\baddr',\eaddr'] \ldotp \npair[n'-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
+  \dominique{TODO: get rid of n'-1 above somehow.}
+
+  Since $a \in [b,e] \subseteq [\baddr',\eaddr']$, we have that $\npair[n'-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
+  we can take $W_M' = W_r' \oplus W_{\var{heap}}' \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})$ with $W_r' = \oplus_{\aaddr \in ([\baddr',\eaddr']\setminus \{\aaddr\})} S(\aaddr)$ and $W_\var{heap}' = \oplus_{r' \in (\dom(\activeReg{\pwheap[W_\var{heap}]})\setminus \{r\})} R_W(r')$, and get
+  \begin{align*}
+    S(\aaddr) \oplus W_M'
+    &=S(\aaddr) \oplus (W_r' \oplus W_\var{heap}'\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}}))\\
+    &=
+    \oplus_{\aaddr \in [\baddr',\eaddr']} S(\aaddr) \oplus W_{\var{heap}}'\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
+    &=
+    \xi(\xi^{-1}(R_W(r))) \oplus W_\var{heap}'\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
+    &=
+    R_W(r) \oplus W_\var{heap}'\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
+    &=
+    \oplus_{r \in \dom(\activeReg{\pwheap[W_\var{heap}]})} R_W(r)\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
+    &=
+      W_\var{heap} \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
+      &= W_M
+  \end{align*}
+
+  Additionally, if $\npair{(b,e)}\in\writeCond{l,W}$, then
+  we get an $r' \in \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$
+  and $[\baddr',\eaddr'] \supseteq [\baddr,\eaddr]$ such that
+  $\pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur}$ and
+  $\pwheap(r) \text{ is address-stratified}$. Because $W \oplus W_M$ is defined, it follows that also 
+  $\pwheap[W_\var{heap}](r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur}$ and
+  $\pwheap[W_\var{heap}](r) \text{ is address-stratified}$.
+  \dominique{TODO: get rid of n-1 above.}
+
+  It follows that $r = r'$ because $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = [\baddr',\eaddr'] = \dom(R_\ms(r').2) = \dom(R_\ms(r').1)$ and all the $R_\ms(r).1$ and $R_\ms(r).2$ are disjoint.
+
+  We have that $\npair[n']{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \;
+  \pwheap[W_\var{heap}](r).s \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur}$.
+
+  From this, it follows that $\memSatHeap{\ms_S\update{a}{0},\ms_\var{T,heap}\update{a}{0}}{W_r' \oplus W_\var{heap}'}$ and finally $\memSat{(\ms_S\update{a}{0},\stk,\ms_\stk,\ms_T\update{a}{0})}{W_M'}$.
+\end{proof}
 
@@
 \begin{enumerate}
-\item $\npair[n']{[(c,c)]} \in \lrp(W)$
+\item $\npair[n']{[(c,c)]} \in \lrp(W)$ \dominique{Weird that this is an assumption here.
+Shouldn't this be a proof obligation?
+Or rather: I think this should not be here at all but in the value relation instead?
+Remove this from the definition of $\lre$?}
 \item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
@@
 \begin{enumerate}
-\item $\txjmp{r_1}{r_2}$
+\item $\tfail$
+\item $\thalt$
+\item $\tjmp{\trg{r_1}}$
+\item $\tjnz{\trg{r_1}}{\trg{\rn}}$
+\item $\tisptr{\trg{r_1}}{\trg{r_2}}$
+\item $\tgeta{\trg{r_1}}{\trg{r_2}}$
+\item $\tgetb{\trg{r_1}}{\trg{r_2}}$
+\item $\tgetp{\trg{r_1}}{\trg{r_2}}$
+\item $\tgetlin{\trg{r_1}}{\trg{r_2}}$
+\item $\tmove{\trg{r_1}}{\trg{\rn}}$
+\item $\tstore{\trg{r_1}}{\trg{r_2}}$
+\item $\tload{\trg{r_1}}{\trg{r_2}}$
+\item $\tcca{\trg{r_1}}{\trg{\rn}}$
+\item $\trestrict{\trg{r_1}}{\trg{r_2}}{\trg{\rn}}$
+\item $\tlt{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$
+\item $\tplus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$
+\item $\tminus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$
+\item $\tsetatob{\trg{r_1}}{\trg{r_2}}$
+\item $\txjmp{\trg{r_1}}{\trg{r_1}}$
+\item $\tcseal{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$
+\item $\tsplice{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$
 \item $\scall[o]{}{r_1}{r_2}$
 \end{enumerate}
+
+\paragraph{Case: $\protect\tfail$}
+Contradiction, because then $\Phi_S \step \failed \not\nstep[i-1] \halted$.
+
+\paragraph{Case: $\protect\thalt$}
+
+\paragraph{Case: $\protect\tjmp{\trg{r}}$}
+
+\paragraph{Case: $\protect\tjnz{\trg{r}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tisptr{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tgeta{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tgetb{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tgete{\trg{r}}{\trg{r}}\mid$}
+
+\paragraph{Case: $\protect\tgetlin{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tmove{\trg{r}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tstore{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tload{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tcca{\trg{r}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\trestrict{\trg{r}}{\trg{r}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tsetatob{\trg{r}}{\trg{r}}$}
 
@@
 
+\paragraph{Case: $\protect\tcseal{\trg{r}}{\trg{r}}{\trg{r}}$}
+
+\paragraph{Case: $\protect\tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}}$}
+
+\paragraph{Case: $\protect\tsplice{\trg{r}}{\trg{r}}{\trg{r}}$}
+
 \subsubsection*{Assuming target machine terminates}

### 2018-03-07 fully-abstract-comp/technical-report.tex
Fixes and typos in memory satisfaction.
Small rephrasing.
Rename heap relation for standard region to something recognizable.
Some work on a lemma about the read and write condition. Add some remarks about problems with spurious later modalities.

@@
     \update{\mem.a}{\scall[0]{o}{r_1}{r_2}}\\
-    \update{\mem.a}{\scall[1]{o}{r_1}{r_2}}\\
+    \update{\mem.a+1}{\scall[1]{o}{r_1}{r_2}}\\
     \dots\\
-    \update{\mem.a}{\scall[\calllen-1]{o}{r_1}{r_2}}\\
+    \update{\mem.a+\calllen-1}{\scall[\calllen-1]{o}{r_1}{r_2}}\\
   \end{array}
@@
 
+\begin{lemma}
+  $\oplus$ is associative and commutative.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
 \subsection{Future world}
@@
       \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
-      W = W_{\var{stack}} \uplus W_{\var{free\_stack}} \uplus W_{\var{heap}} \wedge\\
+      W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
       \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
@@
       \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
+      \quad \ms_0 \uplus \dots \uplus \ms_m = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
       \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World \ldotp \\
@@
     \begin{array}{l}
-      W_\var{heap} = \pwheap \wedge \\
-      \exists R_\ms : \dom(\activeReg{W_\var{heap}}) \fun \MemSeg \times \MemSeg \wedge \\
-      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{heap}})} \pi_2(R_\ms(r)) \wedge \\
-      \quad \ms = \biguplus_{r \in \dom(\activeReg{W_\var{heap}})} \pi_1(R_\ms(r)) \wedge \\
-      \quad \exists R_W : \dom(\activeReg{W_\var{heap}}) \fun \World\ldotp\\
-      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{heap}})} R_W(r) \wedge\\
-      \qquad \forall r \in \activeReg{W_\var{heap}}, n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{heap}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r))
+      \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
+      \qquad W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
+      \qquad \forall r \in \activeReg{\pwheap}, n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \pwheap(r).s \; \xi^{-1}(R_W(r))
     \end{array}
@@
       \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \forall W_\lrr, W_\lrm \ldotp \\
+      \quad \forall W_\lrrs , W_\lrm \ldotp \\
       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
       \qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\\
-      \qquad W \oplus W_\lrr \oplus W_\lrm \\
+      \qquad W \oplus W_\lrrs \oplus W_\lrm \\
       \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
@@
       \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \forall W_\lrr, W_\lrm \ldotp \\
+      \quad \forall W_\lrrs , W_\lrm \ldotp \\
       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrr) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
       \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
-      \qquad W \oplus W_\lrr \oplus W_\lrm\\
+      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
       \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
@@
 \[
-  \lro = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+  \lrol = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
@@
       \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
-      \quad \tand\\
+    \end{array}
+\right\}
+\]
+\[
+  \lror = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
       \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term
@@
 \subsection{Permission based conditions}
+\lau{28-02-2018: These definitions need to be updated to fit with the new worlds}.
 \[
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad W(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
+      \qquad \pwheap(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
     \end{array}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad W(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
+      \qquad \pwfree(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
     \end{array}
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, W}, \ms \ldotp\\
+      \exists r \in \addressable{\lin, \pwheap}, \ms \ldotp\\
       \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
-      \quad W(r) \nequal \stareg[(\ms,\ms)]{\pur}
+      \quad \pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}
     \end{array}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad W(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
-      \qquad W(r) \text{ is address-stratified}
+      \qquad \pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
+      \qquad \pwheap(r) \text{ is address-stratified}
     \end{array}
@@
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
+\dominique{Why $n-1$ above?
+There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr']}{\pur}$.
+Also: readCond doesn't have the $n-1$.}
 \[
@@
 \[
-  \stdreg{A}{v} \defeq (v,*,=,=,H_A)
+  \stdreg{A}{v} \defeq (v,*,=,=,H_A^{\mathrm{std}})
 \]
@@
 \[
-  H_A \; s \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
+  H_A^{\mathrm{std}} \; * \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
     \begin{array}{l}
       \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
-      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms_S)} S(\aaddr) \wedge\\
+      \exists S : A \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
       \quad \forall \aaddr \in A \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
@@
 \lau{01-11-2017: TODO look into what kind of world separation should happen here.}
+\dominique{7-3-2018: is it necessary to have $n-1$ above? there is already a quantification over $n'<n$ in the memory satisfaction relation.}
 

### 2018-03-07 fully-abstract-comp/ftlr-proof-cases.tex
Some work on FTLR proof.

@@
 \begin{lemma}
+  \label{lem:readcond-writecond-work}
   If
@@
 \[
-  \npair{(b,e)}\in \readCond{\lin,W}
+  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
 \]
@@
 \item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
-\item $W \oplus W_R \oplus W_M$ is well-defined.
+\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
 \end{enumerate}
@@
 \subsubsection*{Assuming source machine terminates}
-First show the former of the two, assume $\Phi_S\term[i]$. We now consider the first step of $\Phi_S \term[i]$\footnote{if $i=0$, then we have a contradiction with $\Phi_S\term[i]$.}. 
+First show the former of the two, assume $\Phi_S\term[i]$.
+Since $\Phi_S \neq \halted$, we have that $1 \leq i \leq n'$.
+
+We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$.
+From Lemma~\ref{lem:readcond-writecond-work}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+
+Since $\Phi_S \neq \halted$, this means that it must take a next execution step and also that $\Phi_S \not\step\failed$. 
+This is only possible if one of the following holds:
+\begin{itemize}
+\item $\callCond{\Phi_S,r_1,r_2,\aaddr}$ and
+  $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$ and 
+  $\exec{\Phi_S(\pcreg)}$.
+
+  In this case, we have for all $i = 0..\calllen-1$ that $\npair[n']{(\ms_S(\aaddr+i),\ms_T(\aaddr+i))} \in \lrv(W')$ for some $W'$, and also $\ms_S(\aaddr+i) \in \ints$, so by definition of $\lrv$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i)$.
+\item ($\src{(\neg\callCond{\Phi_S,r_1,r_2,\aaddr}}$ or $\src{\eaddr < \aaddr + \calllen-1) \tand}$) and
+  $\withinBounds{\Phi_S(\pcreg)}$ and $\exec{\Phi_S(\pcreg)}$.
+
+  In this case, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(W')$ for some $W'$, and also $\ms_S(\aaddr) \in \ints$ (since $\Phi_S \not\step\failed$), so by definition of $\lrv$, we have that $\ms_T(\aaddr) = \ms_S(\aaddr)$.
+\end{itemize}
+
+We now consider the first step of $\Phi_S \term[i]$ (we have already mentioned that $i \geq 1$).
 The possible instructions that could have been executed are:
@@
 \paragraph{Case: $\protect\thalt$}
+In this case $\Phi_S \step \halted$.
+We've seen above that $\ms_T(\aaddr) = \ms_S(\aaddr)$, so because $\decInstr{}$ is shared by source and target, we have that $\Phi_T \step\halted$.
 

### 2018-03-08 fully-abstract-comp/technical-report.tex
Some work on FTLR proof.
Leave out rwx pointers from the value relation.
Add a lemma to be proven.
Add a TODO about readXCondition

@@
 The $\decInstr{}$ function must be surjective and injective for all non-$\tfail$ instructions.
+For any $c \in \Caps$, we have that $\decInstr{c} = \tfail$.
 The $\encInstr{}$ function must be injective.
@@
         \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
-        \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
-                                   \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
+        % we are excluding rwx pointers.
+        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
+        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
         \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
@@
 \]
+\dominique{The above needs to require that $\ms$ contains safe values somehow.}
 
@@
       \begin{array}{l}
-        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
+        % \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
         \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}
@@
   \]
-  \dominique{6-3-2018: $\readCond{\lin,W}$ and $\xReadCond{\lin,W}$ seem contradictory.
-    I suspect the latter needs to go? I suspect we don't need to assume the context has no way to modify its own code.}
+  \dominique{6-3-2018: I removed $\readCond{\lin,W}$ and replaced it with $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
 \end{theorem}

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
Add a lemma to be proven.
Fix readXCond to require that ms is safe.

@@
 
+\begin{lemma}[readXCondition implies readCondition]
+  \label{lem:readXCond-readCond}
+  If $\npair{(b,e)}\in\xReadCond{\lin,W}$, then $\npair{(b,e)}\in\readCond{\lin,W}$.
+  \dominique{TODO}
+\end{lemma}
+
 \begin{lemma}[Steps-to termination]
@@
 
-\begin{lemma}
+\begin{lemma}[readCondition works]
   \label{lem:readcond-writecond-work}
@@
 \[
-  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
+  \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
 \]
@@
 
-We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$.
-From Lemma~\ref{lem:readcond-writecond-work}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
+From Lemmas~\ref{lem:readcond-writecond-work} and~\ref{lem:readXCond-readCond}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
 

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
TODO: xreadCond implies readcond.

@@
 
+\begin{lemma}
+  \label{lem:stareg-ms-subeq-stdreg}
+  If
+  \begin{itemize}
+  \item $R \nequal \stareg[(\ms,\ms)]{\pur}$
+  \item $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$
+  \item $\dom(\ms) \supseteq [\baddr,\eaddr]$
+  \end{itemize}
+  Then
+  $R \nsubeq \stdreg{\dom(\ms)}{\pur}$
+\end{lemma}
+\begin{proof}
+  \dominique{TODO: define nsubeq and friends first?}
+\end{proof}
+
 \begin{lemma}[readXCondition implies readCondition]
   \label{lem:readXCond-readCond}
-  If $\npair{(b,e)}\in\xReadCond{\lin,W}$, then $\npair{(b,e)}\in\readCond{\lin,W}$.
-  \dominique{TODO}
+  If $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, then $\npair{(\baddr,\eaddr)}\in\readCond{\lin,W}$.
 \end{lemma}
+\begin{proof}
+  From $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, we get by definition an $r \in \addressable{\lin, \pwheap}$ and an $\ms$ with $\dom(\ms) \supseteq [\baddr,\eaddr]$,  $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$ and $\pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}$.
+
+  The result then follows directly using Lemma~\ref{lem:stareg-ms-subeq-stdreg}.
+\end{proof}
 

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
Discharge some easy cases.

@@
 \paragraph{Case: $\protect\tgeta{\trg{r}}{\trg{r}}$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
 \paragraph{Case: $\protect\tgetb{\trg{r}}{\trg{r}}$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
 \paragraph{Case: $\protect\tgete{\trg{r}}{\trg{r}}\mid$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
 \paragraph{Case: $\protect\tgetlin{\trg{r}}{\trg{r}}$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
Some comments, proposed generalized antireduction lemma.

@@
 
-\begin{lemma}[Steps-to termination]
+\begin{lemma}[$\lro$ closed under target language antireduction]
   For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
@@
   \npair{(\Phi_S,\Phi_T)} \in \lro
+\]
+\end{lemma}
+
+\dominique{9-3-2018: the below is a proposed generalisation of the above antireduction lemma  that I expect will hold (based on experience with similar LRs in the past).}
+\begin{lemma}[$\lro$ closed under antireduction (generalised previous lemma)]
+  For all $\Phi_S$, $\Phi_S'$, $\Phi_T$, $\Phi_T'$, $j_S,J_T$, $n$, if
+  \begin{itemize}
+  \item $\Phi_S \steps[j_S] \Phi_S'$
+  \item $\Phi_T \steps[j_T] \Phi_T'$
+  \item $\npair{(\Phi_S',\Phi_T')} \in \lro$
+\end{itemize}
+then
+\[
+  \npair[n+\min(j_S,j_T)]{(\Phi_S,\Phi_T)} \in \lro
 \]
@@
 \subsubsection*{Assuming target machine terminates}
+\dominique{I suspect it would be more economical to not treat this case separately from the above (source machine terminates), but rather prove both results simultaneously by induction on the length of the terminating execution.}
 \paragraph{Case: }

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
added comment

--- 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
@@
 \begin{lemma}
-  \label{lem:stareg-ms-subeq-stdreg}
-  If
-  \begin{itemize}
-  \item $R \nequal \stareg[(\ms,\ms)]{\pur}$
-  \item $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$
-  \item $\dom(\ms) \supseteq [\baddr,\eaddr]$
-  \end{itemize}
-  Then
-  $R \nsubeq \stdreg{\dom(\ms)}{\pur}$
-\end{lemma}
-\begin{proof}
-  \dominique{TODO: define nsubeq and friends first?}
-\end{proof}
-
-\begin{lemma}[readXCondition implies readCondition]
-  \label{lem:readXCond-readCond}
-  If $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, then $\npair{(\baddr,\eaddr)}\in\readCond{\lin,W}$.
-\end{lemma}
-\begin{proof}
-  From $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, we get by definition an $r \in \addressable{\lin, \pwheap}$ and an $\ms$ with $\dom(\ms) \supseteq [\baddr,\eaddr]$,  $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$ and $\pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}$.
-
-  The result then follows directly using Lemma~\ref{lem:stareg-ms-subeq-stdreg}.
-\end{proof}
-
-\begin{lemma}[Steps-to termination]
-  For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
-\[
-  \Phi_T \steps[j] \Phi_T' \text{ and } \npair{(\Phi_S,\Phi_T')} \in \lro,
-\]
-then
-\[
-  \npair{(\Phi_S,\Phi_T)} \in \lro
-\]
-\end{lemma}
-
-\begin{lemma}[readCondition works]
   \label{lem:readcond-writecond-work}
@@
 \[
-  \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
+  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
 \]
@@
 \begin{enumerate}
-\item $\npair[n']{[(c,c)]} \in \lrp(W)$ \dominique{Weird that this is an assumption here.
+\item $\npair[n']{[(c,c)]} \in \lrp(W)$ 
+\dominique{Weird that this is an assumption here.
 Shouldn't this be a proof obligation?
@@
 Remove this from the definition of $\lre$?}
+\lau{$(c,c)$ is not shown or assumed to be in the value relation, so to make sure it respects lineraity, this condition is placed here.
+Further, this does become a proof obligation when you use the FTLR.}
 \item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
@@
 
-We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
-From Lemmas~\ref{lem:readcond-writecond-work} and~\ref{lem:readXCond-readCond}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$.
+From Lemma~\ref{lem:readcond-writecond-work}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
 

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
Fix changes I made since Lau pointed out that xreadcond is intended to be used *together* with readcond.

--- 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
@@
 \end{lemma}
-
-\begin{lemma}
-  \label{lem:stareg-ms-subeq-stdreg}
-  If
-  \begin{itemize}
-  \item $R \nequal \stareg[(\ms,\ms)]{\pur}$
-  \item $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$
-  \item $\dom(\ms) \supseteq [\baddr,\eaddr]$
-  \end{itemize}
-  Then
-  $R \nsubeq \stdreg{\dom(\ms)}{\pur}$
-\end{lemma}
-\begin{proof}
-  \dominique{TODO: define nsubeq and friends first?}
-\end{proof}
-
-\begin{lemma}[readXCondition implies readCondition]
-  \label{lem:readXCond-readCond}
-  If $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, then $\npair{(\baddr,\eaddr)}\in\readCond{\lin,W}$.
-\end{lemma}
-\begin{proof}
-  From $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, we get by definition an $r \in \addressable{\lin, \pwheap}$ and an $\ms$ with $\dom(\ms) \supseteq [\baddr,\eaddr]$,  $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$ and $\pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}$.
-
-  The result then follows directly using Lemma~\ref{lem:stareg-ms-subeq-stdreg}.
-\end{proof}
 
@@
 \[
+  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
   \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
@@
 
-We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
-From Lemmas~\ref{lem:readcond-writecond-work} and~\ref{lem:readXCond-readCond}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$, $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
+From Lemmas~\ref{lem:readcond-writecond-work}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
 

### 2018-03-09 fully-abstract-comp/technical-report.tex
Fix readXCond to require that ms is safe.
TODO: xreadCond implies readcond.
Fix changes I made since Lau pointed out that xreadcond is intended to be used *together* with readcond.

@@
       \exists r \in \addressable{\lin, \pwheap} \ldotp \\
-      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwheap(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
+      \quad \exists A \supseteq [\baddr,\eaddr] \ldotp \\
+      \qquad \pwheap(r) \nsubeq \stdreg{A}{\pur}
     \end{array}
@@
 \]
-\dominique{The above needs to require that $\ms$ contains safe values somehow.}
+\dominique{9-3-2018: I changed the above to require that $\ms$ contains safe values by requiring that it is in Hstd. Is this reasonable?}
 
@@
       \begin{array}{l}
-        % \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
+        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
         \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}
@@
   \]
-  \dominique{6-3-2018: I removed $\readCond{\lin,W}$ and replaced it with $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
+  \dominique{6-3-2018: I added $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
 \end{theorem}

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
Some trivial changes.

@@
 
-\paragraph{Case: $\protect\tjmp{\trg{r}}$}
+\paragraph{Case: $\protect\tjmp{\trg{r_1}}$}
 
-\paragraph{Case: $\protect\tjnz{\trg{r}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tjnz{\trg{r_1}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tisptr{\trg{r}}{\trg{r}}$}
+\paragraph{Case: $\protect\tisptr{\trg{r_1}}{\trg{r_2}}$}
 
-\paragraph{Case: $\protect\tgeta{\trg{r}}{\trg{r}}$}
-Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
+\paragraph{Case: $\protect\tgeta{\trg{r_1}}{\trg{r_2}}$}
+Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
 
-\paragraph{Case: $\protect\tgetb{\trg{r}}{\trg{r}}$}
-Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
+\paragraph{Case: $\protect\tgetb{\trg{r_1}}{\trg{r_2}}$}
+Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
 
-\paragraph{Case: $\protect\tgete{\trg{r}}{\trg{r}}\mid$}
-Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
+\paragraph{Case: $\protect\tgete{\trg{r_1}}{\trg{r_2}}\mid$}
+Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
 
-\paragraph{Case: $\protect\tgetlin{\trg{r}}{\trg{r}}$}
-Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
+\paragraph{Case: $\protect\tgetlin{\trg{r_1}}{\trg{r_2}}$}
+Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
 
-\paragraph{Case: $\protect\tmove{\trg{r}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tmove{\trg{r_1}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tstore{\trg{r}}{\trg{r}}$}
+\paragraph{Case: $\protect\tstore{\trg{r_1}}{\trg{r_2}}$}
 
-\paragraph{Case: $\protect\tload{\trg{r}}{\trg{r}}$}
+\paragraph{Case: $\protect\tload{\trg{r_1}}{\trg{r_2}}$}
 
-\paragraph{Case: $\protect\tcca{\trg{r}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tcca{\trg{r_1}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\trestrict{\trg{r}}{\trg{r}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\trestrict{\trg{r_1}}{\trg{r_2}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tlt{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tplus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tminus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tsetatob{\trg{r}}{\trg{r}}$}
+\paragraph{Case: $\protect\tsetatob{\trg{r_1}}{\trg{r_2}}$}
 
@@
 
-\paragraph{Case: $\protect\tcseal{\trg{r}}{\trg{r}}{\trg{r}}$}
+\paragraph{Case: $\protect\tcseal{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$}
 
-\paragraph{Case: $\protect\tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tsplit{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tsplice{\trg{r}}{\trg{r}}{\trg{r}}$}
+\paragraph{Case: $\protect\tsplice{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$}
 

### 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
First step for load case.
Preparatory work on load case for FTLR

@@
 
+\begin{definition}[purePart]
+  For any world $W$, we define
+  \begin{align*}
+    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
+    \purePart{W_\var{heap}} &\defeq
+                       \begin{cases}
+                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_\var{heap}(r) = \revoked\\
+                       \end{cases}\\
+    \purePart{W_\var{priv}} &\defeq
+                       \begin{cases}
+                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\sigma)\\
+                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\sigma)\\
+                         (\revoked,\sigma) & \text{if } W_\var{priv}(r) = (\revoked,\sigma)\\
+                       \end{cases}
+    \purePart{W_\var{free}} &\defeq
+                       \begin{cases}
+                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_{\var{free}}(r) = \revoked\\
+                       \end{cases}
+  \end{align*}
+
+\end{definition}
+\begin{lemma}[purePart is pure]
+ For all $W$, we have that $W = W \oplus \purePart{W}$.
+\end{lemma}
+
+\begin{lemma}[non-linear words are pure]
+  If $\npair{(w_1,w_2)} \in \lrv{W}$ and ($\nonLinear{w_1}$ or $\nonLinear{w_2}$), then
+  $\npair{(w_1,w_2)} \in \lrv{\purePart{W}}$.
+\end{lemma}
+
+\begin{lemma}[load from capability works]
+  \label{lem:readcond-writecond-work}
+  If
+  \begin{itemize}
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
+  \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+  \item $\npair{(c,c)}\in\lrv(H_\sigma,W)$
+  \item $W \oplus W_M$ is defined
+  \item $n' < n$
+  \item $w_S = \linCons{\ms_S(a)}$, $w_T = \linCons{\ms_T(a)}$
+  \item $\linConsPerm{\perm}$
+  \end{itemize}
+  Then $\exists W', W_M'\ldotp$
+  Additionally, if
+  \begin{itemize}
+  \item $W_M = W' \oplus W_M'$
+  \item $\memSat{(\ms_S\update{\mem.a}{w_S},\stk,\ms_\stk,\ms_T\update{\mem.a}{w_T})}{W_M'}$
+  \item $\npair{(\ms_S(a),\ms_T(a))} \in \lrv(W')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
 \section{FTLR proof}
@@
 \paragraph{Case: $\protect\tload{\trg{r_1}}{\trg{r_2}}$}
+Since $\Phi_S\not\step\failed$, we have that $\Phi_S \step \Phi_S' \term[i-1]$ and one of the following cases hold: 
+\begin{itemize}
+\item \[\Phi_S' = \updPcAddr{\Phi\update{\mem.a}{w_2}\updReg{r_1}{w}}\]
+  with $\Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ 
+  $\perm \in \readAllowed{}$, $\withinBounds{\Phi(r_2)}$, $w = \Phi.\mem(a)$ and $w_2 = \linCons{w}$ and $\linConsPerm{w,\perm}$
+\item \[\Phi_S' = \updPcAddr{\Phi \update{\ms_\stk.a}{w_2}\updReg{r_1}{w}}\]
+  with $\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $\perm \in \readAllowed{}$, $\withinBounds{\Phi(r_2)}$, $\aaddr \in \dom(\Phi.\ms_\stk)$, $w = \Phi.\ms_\stk(a)$ and $w_2 = \linCons{w}$ and $\linConsPerm{w,\perm}$
+\end{itemize}
 

### 2018-03-11 fully-abstract-comp/ftlr-proof-cases.tex
Some more work on the FTLR cases.

@@
                          (\revoked,\sigma) & \text{if } W_\var{priv}(r) = (\revoked,\sigma)\\
-                       \end{cases}
+                       \end{cases}\\
     \purePart{W_\var{free}} &\defeq
@@
   \end{align*}
+\end{definition}
 
-\end{definition}
 \begin{lemma}[purePart is pure]
@@
 
+\begin{lemma}
+  \label{lem:world-fut-purePart}
+  For all $W$, we have that $W \future \purePart{W}$.
+  \dominique{11-3-2018: Err..
+this doesn't seem to hold as we don't have that $\spatialo \future \spatial$.
+This also means that we don't currently have that $W_1 \oplus W_2 \future W_1$.
+Shouldn't we?
+}
+\end{lemma}
+
 \begin{lemma}[non-linear words are pure]
-  If $\npair{(w_1,w_2)} \in \lrv{W}$ and ($\nonLinear{w_1}$ or $\nonLinear{w_2}$), then
-  $\npair{(w_1,w_2)} \in \lrv{\purePart{W}}$.
+  If $\npair{(w_1,w_2)} \in \lrv(H_\sigma,W)$ and ($\nonLinear{w_1}$ or $\nonLinear{w_2}$), then
+  $\npair{(w_1,w_2)} \in \lrv(H_\sigma,\purePart{W})$.
+  \dominique{11-3-2018: proving this lemma depends on lemma~\ref{lem:world-fut-purePart} for the higher-order cases.}
 \end{lemma}

### 2018-03-12 fully-abstract-comp/ftlr-proof-cases.tex

--- 2018-03-09 fully-abstract-comp/ftlr-proof-cases.tex
@@
 \end{lemma}
+
+\begin{lemma}
+  \label{lem:stareg-ms-subeq-stdreg}
+  If
+  \begin{itemize}
+  \item $R \nequal \stareg[(\ms,\ms)]{\pur}$
+  \item $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$
+  \item $\dom(\ms) \supseteq [\baddr,\eaddr]$
+  \end{itemize}
+  Then
+  $R \nsubeq \stdreg{\dom(\ms)}{\pur}$
+\end{lemma}
+\begin{proof}
+  \dominique{TODO: define nsubeq and friends first?}
+\end{proof}
+
+\begin{lemma}[readXCondition implies readCondition]
+  \label{lem:readXCond-readCond}
+  If $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, then $\npair{(\baddr,\eaddr)}\in\readCond{\lin,W}$.
+\end{lemma}
+\begin{proof}
+  From $\npair{(\baddr,\eaddr)}\in\xReadCond{\lin,W}$, we get by definition an $r \in \addressable{\lin, \pwheap}$ and an $\ms$ with $\dom(\ms) \supseteq [\baddr,\eaddr]$,  $\npair{(\ms,\ms)} \in H^{\mathrm{std}}_{\dom(\ms)} \; * \; W$ and $\pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}$.
+
+  The result then follows directly using Lemma~\ref{lem:stareg-ms-subeq-stdreg}.
+\end{proof}
 
@@
 \[
-  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
   \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
@@
 
-We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$, $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
-From Lemmas~\ref{lem:readcond-writecond-work}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
+From Lemmas~\ref{lem:readcond-writecond-work} and~\ref{lem:readXCond-readCond}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
 
@@
 
-\paragraph{Case: $\protect\tjmp{\trg{r_1}}$}
+\paragraph{Case: $\protect\tjmp{\trg{r}}$}
 
-\paragraph{Case: $\protect\tjnz{\trg{r_1}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tjnz{\trg{r}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tisptr{\trg{r_1}}{\trg{r_2}}$}
+\paragraph{Case: $\protect\tisptr{\trg{r}}{\trg{r}}$}
 
-\paragraph{Case: $\protect\tgeta{\trg{r_1}}{\trg{r_2}}$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
+\paragraph{Case: $\protect\tgeta{\trg{r}}{\trg{r}}$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
-\paragraph{Case: $\protect\tgetb{\trg{r_1}}{\trg{r_2}}$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
+\paragraph{Case: $\protect\tgetb{\trg{r}}{\trg{r}}$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
-\paragraph{Case: $\protect\tgete{\trg{r_1}}{\trg{r_2}}\mid$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
+\paragraph{Case: $\protect\tgete{\trg{r}}{\trg{r}}\mid$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
-\paragraph{Case: $\protect\tgetlin{\trg{r_1}}{\trg{r_2}}$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
+\paragraph{Case: $\protect\tgetlin{\trg{r}}{\trg{r}}$}
+Similar to the case for $\tisptr{\trg{r}}{\trg{r}}$.
 
-\paragraph{Case: $\protect\tmove{\trg{r_1}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tmove{\trg{r}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tstore{\trg{r_1}}{\trg{r_2}}$}
+\paragraph{Case: $\protect\tstore{\trg{r}}{\trg{r}}$}
 
-\paragraph{Case: $\protect\tload{\trg{r_1}}{\trg{r_2}}$}
+\paragraph{Case: $\protect\tload{\trg{r}}{\trg{r}}$}
 
-\paragraph{Case: $\protect\tcca{\trg{r_1}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tcca{\trg{r}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\trestrict{\trg{r_1}}{\trg{r_2}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\trestrict{\trg{r}}{\trg{r}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tlt{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tplus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tminus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tsetatob{\trg{r_1}}{\trg{r_2}}$}
+\paragraph{Case: $\protect\tsetatob{\trg{r}}{\trg{r}}$}
 
@@
 
-\paragraph{Case: $\protect\tcseal{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$}
+\paragraph{Case: $\protect\tcseal{\trg{r}}{\trg{r}}{\trg{r}}$}
 
-\paragraph{Case: $\protect\tsplit{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}{\trg{\rn}}$}
+\paragraph{Case: $\protect\tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}}$}
 
-\paragraph{Case: $\protect\tsplice{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$}
+\paragraph{Case: $\protect\tsplice{\trg{r}}{\trg{r}}{\trg{r}}$}
 

### 2018-03-12 fully-abstract-comp/ftlr-proof-cases.tex
typo and strengthen anti-reduction lemma
changed call semantics, so it does not 'forget' memory not pointed to by the current stack pointer. Also added footnote with explanation in FTLR proof.

@@
 \[
-  \Phi_T \steps[j] \Phi_T' \text{ and } \npair{(\Phi_S,\Phi_T')} \in \lro,
+  \Phi_T \nstep[j] \Phi_T' \text{ and } \npair[n-j]{(\Phi_S,\Phi_T')} \in \lro,
 \]
@@
   \begin{itemize}
-  \item $\Phi_S \steps[j_S] \Phi_S'$
-  \item $\Phi_T \steps[j_T] \Phi_T'$
+  \item $\Phi_S \nstep[j_S] \Phi_S'$
+  \item $\Phi_T \nstep[j_T] \Phi_T'$
   \item $\npair{(\Phi_S',\Phi_T')} \in \lro$
@@
 \[
-  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}
+  \begin{array}{l}
+  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}\\
   \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
+  \end{array}
 \]
@@
 In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution.
-We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}.
+We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{Assuming $\pcreg$ was one of the two registers, then the step relation would reqiure the register to contain an executable capability while the call semantics would require there to be a sealed capability.}\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}.
 By assumption, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
@@
                & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
-               & \update{\mathrm{priv}.y}{\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa}}
+               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa},\opc)}
   \end{array}
 \]
+\lau{12-05-2018: Write what $\opc$ is.}
 where $x$ and $y$ are fresh region names in the respective part of the world.
@@
 and
-$\ms_S,((\dots,\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}) :: \stk),\ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]}$
+$\ms_S,((\opc,\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}) :: \stk),\ms_\stk-\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}$
 

### 2018-03-13 fully-abstract-comp/ftlr-proof-cases.tex
Add a note, fix typo.

@@
 \begin{lemma}[non-linear words are pure]
+  \label{lem:non-linear-pure}
   If $\npair{(w_1,w_2)} \in \lrv(H_\sigma,W)$ and ($\nonLinear{w_1}$ or $\nonLinear{w_2}$), then
@@
   Then $\exists W', W_M'\ldotp$
-  Additionally, if
   \begin{itemize}
@@
   \end{itemize}
+  \dominique{11-3-2018: this lemma depends on Lemma~\ref{lem:non-linear-pure} for the non-linear case.}
 \end{lemma}

### 2018-03-13 fully-abstract-comp/ftlr-proof-cases.tex
Add another lemma about loading from a stack capability.
added the missing simple anti-red. lemma

--- 2018-03-12 fully-abstract-comp/ftlr-proof-cases.tex
@@
 
-\begin{lemma}[$\lro$ closed under target language antireduction]
+\begin{lemma}[$\lror$ closed under target language antireduction]
   For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
 \[
-  \Phi_T \nstep[j] \Phi_T' \text{ and } \npair[n-j]{(\Phi_S,\Phi_T')} \in \lro,
+  \Phi_T \nstep[j] \Phi_T' \text{ and } \npair[n-j]{(\Phi_S,\Phi_T')} \in \lror,
 \]
@@
 \[
-  \npair{(\Phi_S,\Phi_T)} \in \lro
+  \npair{(\Phi_S,\Phi_T)} \in \lror
+\]
+\end{lemma}
+
+\begin{lemma}[$\lrol$ closed under target language antireduction]
+  For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
+\[
+  \Phi_S \nstep[j] \Phi_S' \text{ and } \npair[n-j]{(\Phi_S',\Phi_T)} \in \lrol,
+\]
+then
+\[
+  \npair{(\Phi_S,\Phi_T)} \in \lrol
 \]

### 2018-03-13 fully-abstract-comp/ftlr-proof-cases.tex

@@
 
-\begin{lemma}[load from capability works]
+\begin{lemma}[load from regular capability works]
   \label{lem:readcond-writecond-work}
@@
   \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-  \item $\npair{(c,c)}\in\lrv(H_\sigma,W)$
+  \item $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$
+  \item $\perm \in \readAllowed{}$, $\perm' \in \readAllowed{}$
+  \item $\npair{(c,c')}\in\lrv(H_\sigma,W)$
   \item $W \oplus W_M$ is defined
   \item $n' < n$
-  \item $w_S = \linCons{\ms_S(a)}$, $w_T = \linCons{\ms_T(a)}$
-  \item $\linConsPerm{\perm}$
+  \item $w_S = \linCons{\ms_S(\aaddr)}$, $w_T = \linCons{\ms_T(\aaddr')}$
+  \item $\linConsPerm{\perm}{\ms_S(\aaddr)}$, $\linConsPerm{\perm}{\ms_T(\aaddr')}$
   \end{itemize}
   Then $\exists W', W_M'\ldotp$
-  Additionally, if
   \begin{itemize}
   \item $W_M = W' \oplus W_M'$
-  \item $\memSat{(\ms_S\update{\mem.a}{w_S},\stk,\ms_\stk,\ms_T\update{\mem.a}{w_T})}{W_M'}$
-  \item $\npair{(\ms_S(a),\ms_T(a))} \in \lrv(W')$
+  \item $\memSat{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$
+  \item $\npair{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$
   \end{itemize}
+  \dominique{11-3-2018: this lemma depends on Lemma~\ref{lem:non-linear-pure} for the non-linear case.}
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}[load from stack capability works]
+  \label{lem:readcond-writecond-work}
+  If
+  \begin{itemize}
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
+  \item $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$
+  \item $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$
+  \item $\perm \in \readAllowed{}$
+  \item $\npair{(c,c')}\in\lrv(H_\sigma,W)$
+  \item $W \oplus W_M$ is defined
+  \item $n' < n$
+  \item $w_S = \linCons{\ms_S(\aaddr)}$, $w_T = \linCons{\ms_T(\aaddr')}$
+  \item $\linConsPerm{\perm}{\ms_S(\aaddr)}$, $\linConsPerm{\perm'}{\ms_T(\aaddr')}$
+  \end{itemize}
+  Then $\exists W', W_M'\ldotp$
+  \begin{itemize}
+  \item $W_M = W' \oplus W_M'$
+  \item $\memSat{(\ms_S,\stk,\ms_\stk\update{\aaddr}{w_S},\ms_T\update{\aaddr'}{w_T})}{W_M'}$
+  \item $\npair{(\ms_\stk(a),\ms_T(a))} \in \lrv(W')$
+  \end{itemize}
+  \dominique{11-3-2018: this lemma depends on Lemma~\ref{lem:non-linear-pure} for the non-linear case.}
 \end{lemma}

### 2018-03-13 fully-abstract-comp/technical-report.tex
Problem in semantics of load.
Fix copy-paste error.
Fix load semantics to check writability of the capability through which a linear capability is being loaded.
changed call semantics, so it does not 'forget' memory not pointed to by the current stack pointer. Also added footnote with explanation in FTLR proof.
Add a note.

@@
 \[
-  \isLinear{\cb} \defeq
+  \isLinear{c} \defeq
   \begin{cases}
@@
     \begin{array}[t]{l}
-      \cb = ((\_,\linear),\_,\_,\_) \tor\\
-      \cb = \sealed{\_,\cb'} \tand \isLinear{\cb'} 
+      c = ((\_,\linear),\_,\_,\_) \tor\\
+      c = \sealed{\_,\cb'} \tand \isLinear{\cb'} 
     \end{array}\\
@@
     \begin{array}[t]{l}
-      \cb = \stkptr{\_,\_,\_,\_}\\
-      \cb = \retptrd(\_,\_)\\
-      \cb = \retptrd(\_,\linear,\_,\_,\_)
+      c = \stkptr{\_,\_,\_,\_}\\
+      c = \retptrd(\_,\_)\\
+      c = \retptrc(\_,\linear,\_,\_,\_)
     \end{array}\right.\\
@@
     w & \totherwise
+  \end{cases}
+\]
+\[
+  \linConsPerm{\perm}{w} \defeq
+  \begin{cases}
+    \perm \in \writeAllowed{} & \isLinear{w} \\
+    \mathrm{true} & \totherwise
   \end{cases}
@@
                                       w = \Phi.\mem(a) \tand \\
-                                      w_2 = \linCons{w}
+                                      w_2 = \linCons{w} \tand
+                                      \linConsPerm{\perm}{w}
                                     \end{array}
@@
                                       w = \Phi.\ms_\stk(a) \tand \\
-                                      w_2 = \linCons{w}
+                                      w_2 = \linCons{w} \tand
+                                      \linConsPerm{\perm}{w}
                                     \end{array}\right.
@@
 \end{align*} 
+\dominique{9-3-2018: $\linCons$ should only be allowed to modify the memory if the capability has write permission!}
 
@@
       \begin{array}[t]{l}
-        w_1 = \linCons{c_1}
         c_1 = \retptrc(\sigma_1,\_) \tand \\
@@
                                                    \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} \tand\\
-                                                   \ms_{\stk,\var{rest}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk - 1]} \tand \\
+                                                   \ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} \tand \\
                                                    c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
@@
 And if we do not check it as part of the call implementation, do we then want to match the call implementation better?}
+\dominique{13-3-2018: re the above note: the above should be mitigated by Definition~\ref{def:check-stack-addr-before-call}.
+Adding a check is in principle not necessary, since the check would only protect trusted code from itself, rather than the adversary.}
+\dominique{13-3-2018: where does $o$ come from above in the calculation of $\sigma_?$?}
 
@@
         % we are excluding rwx pointers.
+        \perm \neq \rwx \\
         % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\

### 2018-03-14 fully-abstract-comp/ftlr-proof-cases.tex

@@
 \]
-Let $i \leq n'$ be given \footnote{ignoring the frame...} and show
+That is, let $i \leq n'$ be given \footnote{ignoring the frame...} and show
 \[
@@
 We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{Assuming $\pcreg$ was one of the two registers, then the step relation would reqiure the register to contain an executable capability while the call semantics would require there to be a sealed capability.}\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}.
-By assumption, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
+By assumption \ref{item:reg-ass}, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
 \[
@@
 \[
-  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}
+  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}.
 \]
@@
 To this end, we need to know a bit about $W_{R,1}$.
-From the operational semantics, we know $\Phi_S(\rstk) = \stkptr{\perm,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$.
-From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $W_{R,\stk} \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
+From the operational semantics and in particular the semantics of $\scall{}{}{}{}$, we know $\Phi_S(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$.
+From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree{W_{R,\stk}}(r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
 Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}.
-As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $W_{R,1}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
+As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $\pwfree{W_{R,1}}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
 We pick $W'$ to be
@@
                & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
-               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk']}{\spa},\opc)}
+               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk]}{\spa},\opc)}\\
+               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\addr_\stk']}{\spa}}
   \end{array}
@@
 \lau{12-05-2018: Write what $\opc$ is.}
-where $x$ and $y$ are fresh region names in the respective part of the world.
+where $x$, $y$, and $z$ are fresh region names in the respective part of the world.
 To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t.
 $W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$.
-Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$).
+Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$, so the new region at $y$ is non-empty).
 Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get
@@
 \[
-  \npair[n'']{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1}' \oplus W_{R,2}')
+  \npair[n'']{(\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2))} \in \lrexj(W_{R,1}' \oplus W_{R,2}')
 \]
@@
 \]
-We have to argue (1) that this register-file pair is in the register-file relation for a world $W_R'$ with the same ownership as $W_R$ and regions $z$ and $y$ are owned, and (2) the memory-pair satisfies the memory satisfaction relation for a world $W_M'$ (i.e.
+We have to argue (1) that this register-file pair is in the register-file relation for a world $W_R'$ with the same ownership as $W_R$ and regions $x$ and $y$ are owned, and (2) the memory-pair satisfies the memory satisfaction relation for a world $W_M'$ (i.e.
 with the same ownership as $W_M$).
@@
 Let's start with (2).
-Here we use that we took a standard region and replaced by two standard regions.
-This means that when we split up the memory, we can reuse the same splitting og memory and world (w.r.t memory) except for the memory segment that previously satsfied the standard region in question.
-This memory segment, we split in two, so each piece fit with one of the new regions.
+Here we use that we took a standard region and replaced by three standard regions.
+This means that when we split up the memory, we can reuse the same splitting of memory and world (w.r.t memory) except for the memory segment that previously satsfied the standard region in question.
+This memory segment, we split in three, so each piece fit with one of the new regions.
 
@@
 
-For $\rstk$, we make sure the $z$ region is owned in the world we use.
+For $\rstk$, we make sure the $x$ region is owned in the world we use.
 Now we use that we had a standard region governing the stack to get that everything on there is in the value relation.

### 2018-03-14 fully-abstract-comp/technical-report.tex
started a list of compilation assumptions for the FTLR.
Spatial regions can evolve into spatial_owned ones.
Add the notion of the purePart of a world to the main file.

@@
 %
-The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
+The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$.
+Also: both future region relations allow $\spatial$ regions to become $\spatialo$, which expresses that our system is affine, rather than linear.
 \begin{mathpar}
@@
             { \revoked \privft \revoked }
+            \and
+    \inferrule{ }
+            { \spatialo \privft \spatial }
 \and
@@
             \and
+    \inferrule{ }
+            { \spatialo \pubft \spatial }
+\and
   \inferrule{ (s,s') \in \phi_\pub }
@@
 \end{mathpar}
+
+\begin{definition}[The pure part of a world]
+  For any world $W$, we define
+  \begin{align*}
+    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
+    \purePart{W_\var{heap}} &\defeq
+                       \begin{cases}
+                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_\var{heap}(r) = \revoked\\
+                       \end{cases}\\
+    \purePart{W_\var{priv}} &\defeq
+                       \begin{cases}
+                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\sigma)\\
+                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\sigma)\\
+                         (\revoked,\sigma) & \text{if } W_\var{priv}(r) = (\revoked,\sigma)\\
+                       \end{cases}\\
+    \purePart{W_\var{free}} &\defeq
+                       \begin{cases}
+                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_{\var{free}}(r) = \revoked\\
+                       \end{cases}
+  \end{align*}
+\end{definition}
+
+\begin{lemma}[purePart is duplicable]
+  \label{lem:purePart-duplicable}
+ For all $W$, we have that $W = W \oplus \purePart{W}$.
+\end{lemma}
+
+\begin{lemma}[Increasing authority is the future]
+  \label{lem:oplus-future}
+  For all $W$, we have that:
+  $W_1 \oplus W_2 \future W_1$
+\end{lemma}
+
+\begin{lemma}[Purity is a thing of the past]
+  \label{lem:world-fut-purePart}
+  For all $W$, we have that $W \future \purePart{W}$.
+\end{lemma}
+\begin{proof}
+  Direct consequence of Lemmas~\ref{lem:purePart-duplicable}
+  and~\ref{lem:oplus-future}.
+\end{proof}
 

### 2018-03-14 fully-abstract-comp/ftlr-proof-cases.tex

@@
 \end{proof}
-
-\begin{definition}[purePart]
-  For any world $W$, we define
-  \begin{align*}
-    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
-    \purePart{W_\var{heap}} &\defeq
-                       \begin{cases}
-                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (pure,\var{sm})\\
-                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
-                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
-                         \revoked & \text{if } W_\var{heap}(r) = \revoked\\
-                       \end{cases}\\
-    \purePart{W_\var{priv}} &\defeq
-                       \begin{cases}
-                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\sigma)\\
-                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\sigma)\\
-                         (\revoked,\sigma) & \text{if } W_\var{priv}(r) = (\revoked,\sigma)\\
-                       \end{cases}\\
-    \purePart{W_\var{free}} &\defeq
-                       \begin{cases}
-                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (pure,\var{sm})\\
-                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\
-                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
-                         \revoked & \text{if } W_{\var{free}}(r) = \revoked\\
-                       \end{cases}
-  \end{align*}
-\end{definition}
-
-\begin{lemma}[purePart is pure]
- For all $W$, we have that $W = W \oplus \purePart{W}$.
-\end{lemma}
-
-\begin{lemma}
-  \label{lem:world-fut-purePart}
-  For all $W$, we have that $W \future \purePart{W}$.
-  \dominique{11-3-2018: Err..
-this doesn't seem to hold as we don't have that $\spatialo \future \spatial$.
-This also means that we don't currently have that $W_1 \oplus W_2 \future W_1$.
-Shouldn't we?
-}
-\end{lemma}
 

### 2018-03-16 fully-abstract-comp/ftlr-proof-cases.tex
Some comments and typo fixes in call case.

@@
 In this section, I have listed lemmas that seem to be necessary for the FTLR proof.
-\begin{lemma}[$\lrv$ mono]
-  For all $n$, $w_1$, $w_2$, and $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$.
+\begin{lemma}[non-expansiveness of relations]
+  \label{lem:non-expansive}
+  If $n' \leq n$, then
+  \begin{itemize}
+  \item If $\npair{(b,e)}\readCond{\lin}{W}$, then $\npair[n']{(b,e)}\readCond{\lin}{W}$.
+  \end{itemize}
+\end{lemma}
+
+\begin{lemma}[World monotonicity of relations]
+  \label{lem:monotonicity}
+  For all $n$, $W' \future W$, we have that
+  \begin{itemize}
+  \item If $(w_1,w_2) \in H_\sigma~\sigma~W$, then $(w_1,w_2) \in H_\sigma~\sigma~W'$.
+  \item If $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in
+    \lrv(W')$.
+\end{itemize}
 \end{lemma}
@@
 \begin{lemma}[load from regular capability works]
-  \label{lem:readcond-writecond-work}
+  \label{lem:readcond-cap-works}
   If
@@
 \begin{lemma}[load from stack capability works]
-  \label{lem:readcond-writecond-work}
+  \label{lem:load-stack-cap-works}
   If
@@
 
+{\color{DarkGreen}
+\dominique{Proposed proof refactoring below:}
+First, we prove that one of the following holds:
+\begin{itemize}
+\item $\Phi_S \step \failed$ and $\Phi_T \step\failed$
+\item $\Phi_S \step \halted$ and $\Phi_T \step\halted$
+\item All of the following hold:
+  \begin{itemize}
+  \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
+  \item $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$
+  \item $\exec{\Phi_S(\pcreg)}$
+  \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in \ints$.
+  \item TODO: other useful info
+  \end{itemize}
+\item All of the following hold: (includes simple cases: gettype, geta, getb, gete, getp, getl, lt, plus, minus, that can be handled uniformly)
+  \begin{itemize}
+  \item $\Phi_S \step \Phi_S'$
+  \item $\Phi_T \step \Phi_T'$
+  \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r}{z}} \neq \failed$
+  \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r}{z}} \neq \failed$
+  \item $z \in \ints$
+  \end{itemize}
+\item All of the following hold: (includes cap-manipulation cases: move, cca, restrict, seta2b, cseal, split, splice, that can be handled mostly uniformly)
+  \begin{itemize}
+  \item \item $\Phi_S \step \Phi_S'$
+  \item $\Phi_T \step \Phi_T'$
+  \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r_1\cdots r_k}{w_1\cdots w_k}}\neq \failed$
+  \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r_1\cdots r_k}{w_1'\cdots w_k'}} \neq \failed$
+  \item One of the following holds:
+    \begin{itemize}
+    \item (restrict,cca,seta2b) $w_1 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $w_2 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\Phi_S(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\Phi_T(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\perm' \sqsubseteq\perm$,
+      $k = 1$
+      \dominique{TODO: simplify restrict,seta2b to modify in-place}
+    \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $w_2 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\Phi_S(r_1) = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$ and $\Phi_T(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\perm' \sqsubseteq\perm$,
+      $k = 1$
+      \dominique{TODO: simplify restrict,seta2b to modify in-place}
+    \item ...
+      \end{itemize}
+  \end{itemize}
+\item Other cases: store, load, jmp, jnz
+\end{itemize}
+
+Now assume that one of the configurations terminates etc.
+
+}
+
 \subsubsection*{Assuming source machine terminates}
@@
 We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$, $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
-From Lemmas~\ref{lem:readcond-writecond-work}, using $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n']{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+From Lemma~\ref{lem:readcond-writecond-work}, using Lemma~\ref{lem:non-expansive}, $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n'-1]{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
 
-Since $\Phi_S \neq \halted$, this means that it must take a next execution step and also that $\Phi_S \not\step\failed$. 
+Since $\Phi_S \neq \halted$ and $\Phi_S\term[i]$, it must take a next execution step and $\Phi_S \not\step\failed$. 
 This is only possible if one of the following holds:
@@
 To this end, we need to know a bit about $W_{R,1}$.
-From the operational semantics and in particular the semantics of $\scall{}{}{}{}$, we know $\Phi_S(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\Phi_S(\rstk) \in \lrv(W_{R,\stk})$.
-From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree{W_{R,\stk}}(r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
+From the operational semantics and in particular the semantics of $\scall{}{}{}{}$, we know $\Phi_S(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\npair[n']{(\src{\reg_S(\rstk)},\trg{\reg_T(\rstk)})} \in \lrv(W_{R,\stk})$.
+From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree[W_{R,\stk}](r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
 Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}.
-As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $\pwfree{W_{R,1}}(r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
+As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $\pwfree[W_{R,1}](r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
+\dominique{How do we know $r_1 \neq \rstk$? Exclude in syntax?}
 We pick $W'$ to be
@@
                & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk]}{\spa},\opc)}\\
-               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\addr_\stk']}{\spa}}
+               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\eaddr_\stk']}{\spa}}
   \end{array}
 \]
-\lau{12-05-2018: Write what $\opc$ is.}
+\dominique{16-3-2018: shouldn't $W'$ be defined as a modification of $W_{R,1}$ rather than $W$?}
+\lau{12-03-2018: Write what $\opc$ is.}
 where $x$, $y$, and $z$ are fresh region names in the respective part of the world.
@@
 Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$, so the new region at $y$ is non-empty).
-Using world monotonicity of $H_\sigma \; \sigma$\footnote{{\color{red}This means that $H_\sigma$ must be monotone in the world!}}, we get
+Using Lemma~\ref{lem:monotonicity}, we get
 \[
@@
 which we use to get:
+\dominique{instantiate $n''$ to something?}
 \[

### 2018-03-16 fully-abstract-comp/technical-report.tex
Another lemma that's probably useful.
Rewrote the new future region rules to fit with the region setup.
Some comments and typo fixes in call case.

@@
 \end{mathpar}
-where we take "is non-empty and stack region" to mean the following:
+where we take ``is non-empty'' to mean the following:
 
@@
 \and
-    \inferrule{ }
-            { \spatialo \privft \spatial }
+  \inferrule{ (s,s') \in \phi}
+            { (\spatialo,s',\phi_\pub,\phi,H) \privft (\spatial,s,\phi_\pub,\phi,H)  }
             \and
@@
             \and
-    \inferrule{ }
-            { \spatialo \pubft \spatial }
+  \inferrule{ (s,s') \in \phi_\pub}
+            { (\spatialo,s',\phi_\pub,\phi,H) \pubft (\spatial,s,\phi_\pub,\phi,H)  }
             \and
-  \inferrule{ (s,s') \in \phi_\pub }
+  \inferrule{ (s,s') \in \phi_\pub}
             { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
@@
  For all $W$, we have that $W = W \oplus \purePart{W}$.
+\end{lemma}
+
+\begin{lemma}[purePart is monotone]
+  \label{lem:purePart-mono}
+ For all $W'\future W$, we have that $\purePart{W'} \future \purePart{W}$.
 \end{lemma}
@@
 where $H_\sigma$ is a function from seals to the set of pairs of sealable capabilities that can be sealed with $\sigma$.
+\dominique{16-3-2018: $H_\sigma$ will need to be step-indexed if it mentions (directly or indirectly) the value relation, to ensure well-foundedness of definitions.}
 \lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
-      \qquad \pwheap(r) \text{ is address-stratified}
+      \qquad \pwfree(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
+      \qquad \pwfree(r) \text{ is address-stratified}
     \end{array}
@@
       \begin{array}{l}
-        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
-        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}
+        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge \\
+        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \wedge \\
+        \text{the compilation assumptions on $n$, $W$, ...}
       \end{array}
@@
   \dominique{6-3-2018: I added $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
+  \begin{definition}[The compilation assumptions]
+    No call uses $\rtmp{1}$ for jumping.\\
+    ...
+  \end{definition}
+
 \end{theorem}

### 2018-03-18 fully-abstract-comp/technical-report.tex
typo in ft region relation. The spatial and spatial_owned part was the wrong way around.

@@
 \end{mathpar}
-where we take ``is non-empty'' to mean the following:
+where we take "is non-empty and stack region" to mean the following:
 
@@
   \inferrule{ (s,s') \in \phi}
-            { (\spatialo,s',\phi_\pub,\phi,H) \privft (\spatial,s,\phi_\pub,\phi,H)  }
+            { (\spatial,s',\phi_\pub,\phi,H) \privft (\spatialo,s,\phi_\pub,\phi,H)  }
             \and
@@
   \inferrule{ (s,s') \in \phi_\pub}
-            { (\spatialo,s',\phi_\pub,\phi,H) \pubft (\spatial,s,\phi_\pub,\phi,H)  }
+            { (\spatial,s',\phi_\pub,\phi,H) \pubft (\spatialo,s,\phi_\pub,\phi,H)  }
             \and
@@
 where $H_\sigma$ is a function from seals to the set of pairs of sealable capabilities that can be sealed with $\sigma$.
-\dominique{16-3-2018: $H_\sigma$ will need to be step-indexed if it mentions (directly or indirectly) the value relation, to ensure well-foundedness of definitions.}
 \lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwfree(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
-      \qquad \pwfree(r) \text{ is address-stratified}
+      \qquad \pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
+      \qquad \pwheap(r) \text{ is address-stratified}
     \end{array}

### 2018-03-19 fully-abstract-comp/technical-report.tex
Added trusted address set. Require call to be in trusted address set to be executed.

--- 2018-03-16 fully-abstract-comp/technical-report.tex
@@
 \subsection{Operational Semantics}
+The source machine is parameterized with a set of trusted addresses $\ta$.
+$\ta$ are the only addresses from which the $\scall{}{}{}$ will be interpreted.
+The source machine represents a virtual intermediate machine which we use to argue well-bracketedness and local state encapsulation.
+It is not meant to be the machine that the actual code is executed on.
+Further, it is the well-bracketedness and local state encapsulation of the compiled code, not the context we are concerned with.
+On the other hand, the target machine is not parameterized with $\ta$ as all the instructions on the target machine is available to the adversary.
+
 \subsubsection{Notes}
@@
 % Lau: 18-12-2017: I do not think this is true anymore?
-% \item Programs have undefined behavior when passing ret/stk pointers.
+% \item Programs have undefined behavior when passing ret/stk pointers.[\aaddr,\baddr]
 \end{itemize}
@@
                                                   \begin{array}[t]{l}
-                                                    \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
+                                                    \text{if }\Phi(\pcreg) 
+= ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
                                                     \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
+                                                    \lbrack \aaddr ,\baddr \rbrack \subseteq \ta \tand \\
                                                     \baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr \tand \\

### 2018-03-19 fully-abstract-comp/technical-report.tex
Add relation on sealed things to world.

--- 2018-03-16 fully-abstract-comp/technical-report.tex
@@
 where $\RegionName = \nats$.
-\[
+\begin{multline*}
   \Regionh = 
-    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \URel{\MemSeg^2})
-\]
+  \{\pure \} \times \State \times \Rels \times\\
+  (\State \fun \World \monnefun \URel{\MemSeg^2}) \times \\
+  (\State \fun \Seal \parfun \World \monnefun \URel{\Caps \times \SealableCaps})
+\end{multline*}
 and
@@
 Notice that no region allows for both linear and non-linear capabilities to address it.
+Notice also that $\pure$ regions have an additional component that allows them to claim ownership of part of the address space of seals and impose a relational invariant on everything signed with those seals.
 
@@
 \subsubsection{\color{DarkGreen} first thoughts on change to future world relation}
-{\color{DarkGreen}
+% {\color{DarkGreen}
 The public part of the future world relation stays the same:
@@
 
-}
+% }
 
@@
 \lau{14-02-2018: No matter what the memory satisfaction ends up looking like, remember to handle the revoked case.}
-{\color{DarkGreen}
+% {\color{DarkGreen}
 Memory satisfaction for new worlds:
@@
       \qquad \forall r \in \activeReg{\pwheap}, n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \pwheap(r).s \; \xi^{-1}(R_W(r))
+      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \pwheap(r).s \; \xi^{-1}(R_W(r)) \wedge\\
+      \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
+      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}) \text{ is defined }\\
+      \quad \dom(\pwheap(r).H_\sigma \; \pwheap(r).s) = R_\var{seal}(r)
     \end{array}
@@
 \]
-}
+% }
 

### 2018-03-19 fully-abstract-comp/technical-report.tex
change restrict and seta2b instructions to modify a capability in-place, which is easier.
Add another sanity condition to the condition of full abstraction.

@@
       & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
-  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}}{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
+  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
       & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
@@
     \tmove{\rtmp{1}}{\pcreg}\\
-    \tsetatob{\rtmp{1}}{\rtmp{1}}\\
+    \tsetatob{\rtmp{1}}\\
     \tload{\rtmp{1}}{\rtmp{1}}\\
@@
 \begin{align*}
-  \sem{\trestrict{r_1}{r_2}{\rn}}(\Phi) = &
+  \sem{\trestrict{r_1}{\rn}}(\Phi) = &
                                       \begin{cases}
@@
                                           \begin{array}{rl}
-                                          \Phi&\updReg{r_2}{w_2}\\
-                                              &\updReg{r_1}{c}
+                                          \Phi &\updReg{r_1}{c}
                                           \end{array} \right)
@@
                                         \begin{array}[t]{l}
-                                          \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
+                                          \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                           \Phi(\rn) = n \tand\\
                                           \decPerm{n} \sqsubseteq \perm \tand \\
-                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                          w_2 = \linCons{\Phi(r_2)}
+                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr)
                                         \end{array}
@@
                                           \begin{array}{rl}
-                                          \Phi&\updReg{r_2}{0}\\
-                                                    &\updReg{r_1}{c}
+                                          \Phi &\updReg{r_1}{c}
                                           \end{array}\right)\right.
@@
                                         \begin{array}[t]{l}
-                                          \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
+                                          \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                           \Phi(\rn) = n \tand\\
@@
 \begin{align*}
-  \sem{\tsetatob{r_1}{r_2}}(\Phi) = & 
+  \sem{\tsetatob{r_1}}(\Phi) = & 
                                 \begin{cases}
                                   \updPcAddr{}\left(
-                                    \arraycolsep=0pt
-                                    \begin{array}{rl}
-                                    \Phi&\updReg{r_2}{w_2}\\
-                                        &\updReg{r_1}{c}
-                                    \end{array} \right)
+                                    \Phi \updReg{r_1}{c}
+                                    \right)
 &
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
-                                      w_2 = \linCons{\Phi(r_2)}\\
+                                      \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
                                       c = ((\perm,\lin),\baddr,\eaddr,\baddr)
@@
                                   \updPcAddr{}\left(
-                                    \arraycolsep=0pt
-                                    \begin{array}{rl}
-                                    \Phi&\updReg{r_2}{w_2}\\
-                                        &\updReg{r_1}{c}
-                                    \end{array} \right)
+                                    \Phi \updReg{r_1}{c}
+                                    \right)
 &
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
-                                      w_2 = \linCons{\Phi(r_2)}\\
+                                      \Phi(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \tand \\
                                       c = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
@@
                                   \updPcAddr{}\left(
-                                    \arraycolsep=0pt
-                                    \begin{array}{rl}
-                                    \Phi&\updReg{r_2}{0}\\
-                                        &\updReg{r_1}{c}
-                                    \end{array} \right)
+                                    \Phi\updReg{r_1}{c}
+                                    \right)
 \right.
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\_} \tand \\
+                                      \Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\_} \tand \\
                                       c = \stkptr{\perm,\baddr,\eaddr,\baddr}
@@
   \label{def:check-stack-addr-before-call}
-  A configuration $\src{\Phi}$ is said to have reasonable calls in $A$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to have reasonable calls in $\ta$ for $n$ steps iff\\
   for all $i \leq n$ if
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\scall{o}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2'}$ and there exists no $j<i$ s.t.
-  \[
-    \src{\Phi} \nstep[j] \src{\Phi''}
-  \]
-  where $\src{\Phi''}$ points to $\src{\scall{o}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
+  and $\src{\Phi'}$ points to $\src{\scall{o}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2}$
+  % and there exists no $j<i$ s.t.
+  % \[
+  %   \src{\Phi} \nstep[j] \src{\Phi''}
+  % \]
+  % where $\src{\Phi''}$ points to $\src{\scall{o}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
   then
   \[
-    \src{\Phi}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
-  \]
-  and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
+    \src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
+  \]
+  % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
+\end{definition}
+
+\begin{definition}[Use linker seals properly]
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
+  for all $i \leq n$, if
+  \[
+    \src{\Phi} \nstep[i] \src{\Phi'}
+  \]
+  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_2)
+  = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\aaddr \in [\baddr,\eaddr]$, $\aaddr
+  \in \ta$ and $\Phi.\mem(\aaddr) = \seal{\sigma_b,\sigma_e}$, then
+  $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$.
 \end{definition}
@@
   \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
+  \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
   \end{itemize}

### 2018-03-19 fully-abstract-comp/technical-report.tex
Added seal case to value relation
Fix broken reference.

@@
 Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the code and data parts of the sealed return closure.
-Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:source-language}, this has the semantic effect that one would expect.
+Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:syntax}, this has the semantic effect that one would expect.
 Similarly, rather than invoking \sreturn{}, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.

### 2018-03-19 fully-abstract-comp/technical-report.tex
typo
Thoughts on the condition for full abstraction, todo: add a notion of used seals to components.

@@
 
-\begin{definition}[Use linker seals properly]
+\begin{definition}[Use return seals only for call]
   A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
@@
   $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$.
+  \dominique{TODO: this should be allowed to be able to construct closures. Not allowed only for return pointer seals.}
 \end{definition}
 
+\begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
+  for all $i \leq n$, if
+  \[
+    \src{\Phi} \nstep[i] \src{\Phi'} \step \src{\Phi''}
+  \]
+  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and
+  $\src{\Phi}.\reg(r_i) = \seal{\sigma_b,\sigma_e}$ and
+  \dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
+\end{definition}
+
+\dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures).
+That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.}
 \begin{definition}
-  A program $\program$ with memory segment $\ms_\program$ is said to be reasonably compiled if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  A program $\program$ with memory segment $\ms_\program$ and $\dom(\ms_\program) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
   \begin{itemize}
@@
   \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
+  \item $\plug{\context}{\program}$ never passes trusted seals accross the boundary.
   \end{itemize}
@@
   \label{thm:full-abstraction}
-  For programs $\program_1$ and $\program_2$ that are reasonably compiled, we have
+  For reasonable programs $\program_1$ and $\program_2$, we have
   \begin{gather*}

### 2018-03-19 fully-abstract-comp/ftlr-proof-cases.tex
change restrict and seta2b instructions to modify a capability in-place, which is easier.
New treatment of H_sigma as non-global.

@@
   $\npair{(w_1,w_2)} \in \lrv(H_\sigma,\purePart{W})$.
-  \dominique{11-3-2018: proving this lemma depends on lemma~\ref{lem:world-fut-purePart} for the higher-order cases.}
 \end{lemma}
@@
       $k = 1$
-      \dominique{TODO: simplify restrict,seta2b to modify in-place}
     \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $w_2 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\Phi_S(r_1) = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$ and $\Phi_T(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\perm' \sqsubseteq\perm$,
       $k = 1$
-      \dominique{TODO: simplify restrict,seta2b to modify in-place}
     \item ...
@@
 \paragraph{Case: $\scall[o]{}{r_1}{r_2}$}
+\dominique{Adapt to new treatment of $H_\sigma$ as non-global}
 In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution.
@@
 \paragraph{Case: $\protect\txjmp{r_1}{r_2}$} 
+\dominique{Adapt to new treatment of $H_\sigma$ as non-global}
 In the case $\txjmp{r_1}{r_2}$\footnote{$\pcreg$ is a corner case here.

### 2018-03-19 fully-abstract-comp/technical-report.tex
New treatment of H_sigma as non-global.

@@
       \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
-      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}) \text{ is defined }\\
+      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \text{ is defined }\\
       \quad \dom(\pwheap(r).H_\sigma \; \pwheap(r).s) = R_\var{seal}(r)
@@
 
-\lau{TODO: Rewrite the other definitions to pass around $H_\sigma$ (if this is how we want to handle seals.)}
 % \[
@@
 
-where $H_\sigma$ is a function from seals to the set of pairs of sealable capabilities that can be sealed with $\sigma$.
-\dominique{16-3-2018: $H_\sigma$ will need to be step-indexed if it mentions (directly or indirectly) the value relation, to ensure well-foundedness of definitions.}
-\lau{28-11-2017: Do we want the execute condition for $\rwx$ as we are doing write xor execute?}
-
 % \dominique{22-1-2018: the below is an alternative proposal to the above definition.
@@
 \[
-  \lrv(H_\sigma,W) =
+  \lrv(W) =
   \begin{array}[t]{l}
@@
       \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
-      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
-      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
+      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
+      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
+      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
       \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
@@
     \right\}\cup\\
-%
-    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
-    % \right\} \cup \\
-%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    \; \middle| \;
+    \exists r \in \dom(\pwheap) \ldotp \forall \ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand H_\sigma\; s \; \sigma \nequal \lrv
+    \right\} \cup \\
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
@@
       \begin{array}{r l l }
-        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
-        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W} \wedge\\
         % we are excluding rwx pointers.
-        \perm \neq \rwx \\
+        \perm \neq \rwx \wedge \\
         % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\

### 2018-03-19 fully-abstract-comp/technical-report.tex
Fix typo.
First stab at the trusted value relation.
Add notes about illegal source program behavior we need to exclude.

@@
     \; \middle| \;
-    \exists r \in \dom(\pwheap) \ldotp \forall \ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand H_\sigma\; s \; \sigma \nequal \lrv
+    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand H_\sigma\; s \; \sigma \nequal \lrv
     \right\} \cup \\
@@
     \right\}
+  \end{array}
+\]
+
+\[
+  \lrvtrusted(W) =
+  \begin{array}[t]{l}
+    \lrv(W)\cup \\
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    \; \middle| \;
+    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \text{TODO: code region for trusted code}
+    \right\} \cup \\
   \end{array}
@@
   \dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
+  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
+  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}
+  \dominique{TODO: Also not allowed: sealing such a value with a seal that you do not own.}
 \end{definition}

### 2018-03-20 fully-abstract-comp/technical-report.tex
PC-derived code pointers also in trusted value relation.
First write up of code region - still needs some work
Trusted and untrusted versions of lrv and lrr, with superscript parameter.

@@
 \[
-  \lrr(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
     \begin{array}{l}
@@
       \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
-      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
+      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
     \end{array}
@@
 \]
-We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simlpy omit that argument.
+We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
 % DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
@@
     \right\} \cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \begin{array}{l}
+      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \text{TODO: code region for trusted code} \wedge\\
+      \perm \sqsubseteq \rx \wedge [\baddr,\eaddr] \subseteq \dom(\var{code})
+    \end{array}\right\}
   \end{array}
@@
 % \]
+
+\[
+  \codereg{\ta,\sigma_\baddr,\sigma_\eaddr,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigma_\baddr \; \sigma_\eaddr \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigma_\baddr \; \sigma_\eaddr \; \code)
+\]
+
+\[
+  H^\mathrm{code} \; \ta \; \sigma_\baddr \; \sigma_\eaddr \; \code \; s \; \hat{W} = \{\npair{(\code,\code)} | \dom(\code) = [\aaddr,\baddr] \tand \code(a) = \seal{\sigma_\baddr,\sigma_\baddr,\sigma_\eaddr}\}
+\]
+\lau{19-03-2018: Add the link capabilitiy to the above? code in ta?}
+\[
+  H^\mathrm{code}_\sigma \; \ta \; s \; \sigma \; \hat{W} = 
+  \begin{array}[t]{l}
+\left\{
+    \begin{array}{l}
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\sigma,((\rx,\lin),\baddr,\eaddr,\aaddr'+\calllen - 1)),\\((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+      \\
+      \begin{array}{l}
+        \decInstr{\code([a',a' + \calllen]-1)} = \overline{\scall{o}{r_1}{r_2}} \tand \\
+        \aaddr = \aaddr' + \retoffset \tand \\
+        \sigma = \sigma_\baddr + o \tand\\
+        \lbrack \baddr, \eaddr \rbrack = \dom(\code) \subseteq \ta
+      \end{array}
+
+    \end{array}
+      \right\} \uplus \\
+\left\{
+    \begin{array}{l}
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\sigma,(\baddr,\eaddr)),\\((\rx,\linear),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+      \\
+      \begin{array}{l}
+        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, ((\rx,\lin),\baddr',\eaddr',\aaddr')) \tand \\
+        \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
+        \quad \decInstr{\code([a',a' + \calllen]-1)} = \overline{\scall{o}{r_1}{r_2}} \tand \\
+        \quad \aaddr = \aaddr' + \calllen \tand \\
+        \quad \sigma = \sigma_\baddr + o \tand\\
+        \quad \lbrack \baddr', \eaddr' \rbrack = \dom(\code) \subseteq \ta
+      \end{array}
+    \end{array}
+      \right\} \uplus\\
+    \left\{
+    \begin{array}{l}
+\left. \npair{\arraycolsep=0pt\left(\array{l}
+      \sealed{\sigma,\vsc}, \vsc' \endarray\right)} \middle| \right. \\
+      \\
+      \begin{array}{l}
+        \sigma \in [\sigma_\baddr,\sigma_\eaddr] \tand\\
+        (\dom(\code) \cup \ta \neq \emptyset \tand \npair{\vsc,\vsc'} \in \lrv \; W) \tor \\
+        \text{not a call}
+      \end{array}
+    \end{array}
+      \right\}
+  \end{array}
+\]
 

### 2018-03-20 fully-abstract-comp/technical-report.tex
trusted and untrusted expression relations.
changed return tokens to represent capabilities instead of sealed capabilities

@@
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
+    & & & {\sourcecolor{} \; \mid \retptrd(\basebnf,\aendbnf) \mid \retptrc(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf)}\\
+    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}\\ 
     &\rbnf & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
@@
     \src{\baddr} & \src{\text{if $\vsc = \stkptr{\_,\baddr,\_,\_}$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrd(\_,(\baddr,\_))$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\baddr,\_,\_))$}}
+    \src{\baddr} & \src{\text{if $\vsc = \retptrd(\baddr,\_)$}} \\ 
+    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\_,\_,\baddr,\_,\_)$}}
   \end{cases}
@@
     \src{\eaddr} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrd(\_,(\_,\eaddr))$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,(\_,\_,\_,\eaddr,\_))$}}
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrd(\_,\eaddr)$}} \\ 
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,\_,\_,\eaddr,\_)$}}
   \end{cases}
@@
     \src{\linear} & \src{\text{if $\vsc = \stkptr{\_,\_,\eaddr,\_}$}} \\ 
-    \src{\linear} & \src{\text{if $\vsc = \retptrd(\_,(\_,\_))$}} \\ 
-    \src{\lin} & \src{\text{if $\vsc = \retptrc(\_,(\_,\lin,\_,\_,\_))$}}
+    \src{\linear} & \src{\text{if $\vsc = \retptrd(\_,\_)$}} \\ 
+    \src{\lin} & \src{\text{if $\vsc = \retptrc(\_,\lin,\_,\_,\_)$}}
   \end{cases}
@@
 \[
-\encType{}_{\var{src}}(\src{\retptrd(\_,(\_,\_))}) = \encType{}_{\var{src}}(\src{\retptrc(\_,(\_,\_,\_,\_,\_))}) = \encType{}_{\var{trg}}(\sealed{\_,\_})
+\encType{}_{\var{src}}(\src{\retptrd(\_,\_)}) = \encType{}_{\var{src}}(\src{\retptrc(\_,\_,\_,\_,\_)}) = \encType{}_{\var{trg}}(((\_,\_),\_,\_,\_))
 \]
@@
                                      \begin{array}[t]{l}
-                                       \Phi(r_1) = c_1 \tand
-                                       \Phi(r_2) = c_2 \tand\\
+                                       \Phi(r_1) = \sealed{\sigma_1,c_1} \tand
+                                       \Phi(r_2) = \sealed{\sigma_2,c_2} \tand\\
+                                       \sigma_1 = \sigma_2 \tand \\
                                        w_1 = \linCons{c_1} \tand \\
@@
     \begin{array}[t]{rl}
-      \Phi & \updReg{\pcreg}{c_\code}\\
-           & \updReg{\rdata}{c_\data}
+      \Phi & \updReg{\pcreg}{c_1}\\
+           & \updReg{\rdata}{c_2}
     \end{array} &
     \begin{array}[t]{l}
-      c_1 = \sealed{\sigma_1,c_\code} \tand \\
-      c_2 = \sealed{\sigma_2,c_\data} \tand \\
-      \sigma_1 = \sigma_2 \tand \\
+      c_1 \neq \retptrc(\_) \tand \\
+      c_2 \neq \retptrd(\_) \tand \\
       \nonExec{c_\data}
@@
       \begin{array}[t]{l}
-        c_1 = \retptrc(\sigma_1,\_) \tand \\
-        c_2 = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
-        \sigma_1 = \sigma_2 \tand\\
+        c_1 = \retptrc(\_) \tand \\
+        c_2 = \retptrd(\aaddr_\stk,\eaddr_{\stk,\priv}) \tand \\
         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\ 
@@
 \reg.\rretc \mapsto \\
-\hfill \retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))
+\hfill \sealed{\sigma_?,\retptrc(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen)}
 \endarray
 \right]\\
-                                                        &\updReg{\rretd}{\retptrd(\sigma_?,(\aaddr_\stk,\eaddr_\stk))}\\
+                                                        &\updReg{\rretd}{\sealed{\sigma_?,\retptrd(\aaddr_\stk,\eaddr_\stk)}}\\
                                                         &\updReg{\rtmp{1}}{0}
@@
   \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
-  \item $\src{w = \retptrc(\_,\_)}$
+  \item $\src{w = \retptrc(\_)}$
   \item $\src{w = \retptrd(\_,\_)}$
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{c_S},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
-    \begin{array}{l}
-      \npair{(\src{c_S},\sealed{\sigma,\vsc_T})} \in \lrp(W) \wedge \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
+    \begin{array}{l}
+      \npair{(\src{\vsc_S},\vsc_T)} \in \lrp(W) \wedge \\
       \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
-      \stpair[.]{c_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
-      \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
-      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
-      \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
+      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
+      \quad \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
+      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o) 
+      % This case was never used (?) so it is not necessary.
+      % \wedge \\
+      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
     \end{array}
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\sigma,((\rx,\lin),\baddr,\eaddr,\aaddr'+\calllen - 1)),\\((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\sealed{\sigma,\retptrc(\rx,\lin,\baddr,\eaddr,\aaddr'+\calllen - 1)},\\((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
       \\
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\sigma,(\baddr,\eaddr)),\\((\rx,\linear),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\sealed{\sigma,\retptrd(\baddr,\eaddr)},\\((\rx,\linear),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
       \\
@@
 As such, we introduce special token values that play the role of the stack pointer/return pointer and we allow all interactions with them that are possible using instructions like \sstore{}{}, \smove{}{} etc.
-Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the code and data parts of the sealed return closure.
+Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the sealed part of a code and data return closure.
 Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:syntax}, this has the semantic effect that one would expect.
-Similarly, rather than invoking \sreturn{}, one would ccall the $\retptrd$ and $\retptrc$ sealed code/data pair.
+Similarly, rather than invoking \sreturn{}, one would ccall \lau{now xjmp?} the $\retptrd$ and $\retptrc$ sealed code/data pair.
 

### 2018-03-20 fully-abstract-comp/technical-report.tex
Comment out unneeded case in value relation for sealed caps.
Remove obsolete title.

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
 \subsection{Future world}
-\subsubsection{\color{DarkGreen} first thoughts on change to future world relation}
 % {\color{DarkGreen}
@@
 \begin{align*}
-  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+  \lrexjg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
     \begin{array}{l}
@@
       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr (W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
     \right\}\\
-  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
+  \lreg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
     \begin{array}{l}
@@
       \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}}]} \in \lrp(W) \wedge \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr (W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
       \forall W' \future W, W_o, n' < n, \stpair[.]{c_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
-      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o) \wedge \\
-      \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
+      \quad \npair[n']{\src{c_S},\src{c_S'},\sealed{\sigma,\vsc_T},\sealed{\sigma,\vsc_T'}} \in \lrexj(W'\oplus W_o)
+      % 20-3-2018: I expect the next case (commented out below) is never used and hence unneeded.
+      % \wedge \\
+      % \quad \npair[n']{\src{c_S'},\src{c_S},\sealed{\sigma,\vsc_T'},\sealed{\sigma,\vsc_T}} \in \lrexj(W' \oplus W_o)
     \end{array}
@@
 It expresses that any reasonable executable capability is safe to use as the pc capability.
-\dominique{I'm following the example of the previous FTLR by not allowing $\rwx$, but it's not clear to me why that's the right thing to do.}
 \begin{theorem}[FTLR]
   \label{thm:ftlr}
-  \[
-    \begin{array}{l}
-      \forall n,W,\lin,\baddr,\eaddr,\aaddr \ldotp \\
-      \quad \left(
-      \begin{array}{l}
-        \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge \\
-        \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \wedge \\
-        \text{the compilation assumptions on $n$, $W$, ...}
-      \end{array}
-      \right)\\
-      \quad\Rightarrow 
+  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
+  If
+  \begin{itemize}
+  \item $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
+  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
+  \end{itemize}
+  and one of the following sets of requirements holds:
+  \begin{itemize}
+  \item \begin{itemize}
+    \item $[\baddr,\eaddr] \subseteq \ta$
+    \item the compilation assumptions on $n$, $W$, ...
+    \item the full abstraction conditions (difference with the previous point?)
+    \end{itemize}
+  \item
+    \begin{itemize}
+    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
+    \end{itemize}
+  \end{itemize}
+  Then
+  \[
     \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
       ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)
-
-    \end{array}
   \]

### 2018-03-20 fully-abstract-comp/technical-report.tex
Update FTLR to trusted/untrusted expression relation.
typos@

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
 
-The source machine allows for variable length instructions which we utilize for $\scall{o}{}{}$.
+The source machine allows for variable length instructions which we utilise for $\scall{o}{}{}$.
 In other words, when $\scall{o}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction (and be within the range of the pc capability).
@@
 Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
-The argument for having one relation relate pairs of pairs of capabilties and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
+The argument for having one relation relate pairs of pairs of capabilities and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
 \begin{align*}
-  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+  \lrexjg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
     \begin{array}{l}
@@
 \]
-\lau{19-03-2018: Add the link capabilitiy to the above? code in ta?}
+\lau{19-03-2018: Add the link capability to the above? code in ta?}
 \[
@@
   \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
-  \item $\plug{\context}{\program}$ never passes trusted seals accross the boundary.
+  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
   \end{itemize}

### 2018-03-20 fully-abstract-comp/technical-report.tex

@@
 
-The source machine allows for variable length instructions which we utilise for $\scall{o}{}{}$.
+The source machine allows for variable length instructions which we utilize for $\scall{o}{}{}$.
 In other words, when $\scall{o}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction (and be within the range of the pc capability).
@@
 Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
-The argument for having one relation relate pairs of pairs of capabilities and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
+The argument for having one relation relate pairs of pairs of capabilties and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
 \begin{align*}
@@
 \]
-\lau{19-03-2018: Add the link capability to the above? code in ta?}
+\lau{19-03-2018: Add the link capabilitiy to the above? code in ta?}
 \[
@@
     \item $[\baddr,\eaddr] \subseteq \ta$
+    \item $\trust = \trusted$
     \item the compilation assumptions on $n$, $W$, ...
@@
     \item $[\baddr,\eaddr] \mathrel{\#} \ta$
+    \item $\trust = \untrusted$
     \end{itemize}
@@
     \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
-      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lre(W)
+      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lreg{\trust}(W)
   \]
@@
   \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
-  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
+  \item $\plug{\context}{\program}$ never passes trusted seals accross the boundary.
   \end{itemize}

### 2018-03-20 fully-abstract-comp/technical-report.tex
Small update.

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
 Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:syntax}, this has the semantic effect that one would expect.
-Similarly, rather than invoking \sreturn{}, one would ccall \lau{now xjmp?} the $\retptrd$ and $\retptrc$ sealed code/data pair.
+Similarly, rather than invoking \sreturn{}, one would jump to the $\retptrd$ and $\retptrc$ sealed code/data pair.
 

### 2018-03-20 fully-abstract-comp/technical-report.tex
Fix small latex issue: \dom does not take a latex argument.
a couple more changes related to retptrc/d change
updated execCond and sealed capability case in LR to have a linera and normal case
Minor fixes.
Small fix.
minor code region changes
Minor fixes in Hcodesigma.
Require current address of stack pointer to be correct value.
Some changes to the code region.

@@
     \forall a \in \dom(\ms) \ldotp\\
-    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom{\ms}
+    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom(\ms)
   \end{multline*}
@@
   \[
-    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom{\ms \cup \ms_s}
+    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom(\ms \cup \ms_s)
   \]
@@
     \forall a \in \dom(\ms) \ldotp\\
-    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom{\ms \cup \ms_s}
+    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom(\ms \cup \ms_s)
   \end{multline*}
@@
   (\State \fun \World \monnefun \URel{\MemSeg^2}) \times \\
-  (\State \fun \Seal \parfun \World \monnefun \URel{\Caps \times \SealableCaps})
+  (\State \fun \Seal \parfun \World \monnefun \URel{\SealableCaps \times \SealableCaps})
 \end{multline*}
@@
       \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
-      \quad \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
-      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o) 
+      \quad (\isLinear{\vsc_1} \Rightarrow \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
+      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+      \quad (\nonLinear{\vsc_1} \Rightarrow \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
+      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
+
       % This case was never used (?) so it is not necessary.
@@
     \begin{array}{l}
-      \forall n' < n, W' \future W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr], W'' \ldotp\\
-      \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}]} \in \lrp(W' \oplus W'') \Rightarrow \\
-      \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}} \in \lre(W' \oplus W'')
+      \forall W_l \ldotp\\
+      \quad ((W_l = \purePart{W} \wedge \lin = \normal) \vee (W_l = W \wedge \lin = \linear)) \wedge\\
+      \quad \forall n' < n, W' \future W_l, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
+      \quad \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}]} \in \lrp(W') \Rightarrow \\
+      \qquad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}} \in \lre(W')
     \end{array}
@@
 \[
-  \codereg{\ta,\sigma_\baddr,\sigma_\eaddr,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigma_\baddr \; \sigma_\eaddr \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigma_\baddr \; \sigma_\eaddr \; \code)
-\]
-
-\[
-  H^\mathrm{code} \; \ta \; \sigma_\baddr \; \sigma_\eaddr \; \code \; s \; \hat{W} = \{\npair{(\code,\code)} | \dom(\code) = [\aaddr,\baddr] \tand \code(a) = \seal{\sigma_\baddr,\sigma_\baddr,\sigma_\eaddr}\}
+  \codereg{\ta,\overline{\sigma_{\mathrm{ret}}},\overline{\sigma_{\mathrm{clos}}},\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code,H^\mathrm{code}_\sigma \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code)
+\]
+
+\[
+  H^\mathrm{code} \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code \; s \; \hat{W} = \left\{
+    \npair{(\code,\code)} \mid
+  \begin{multlined}
+    \dom(\code) = [\aaddr,\baddr] \tand (\overline{\sigma_{\mathrm{ret}}} \cup \overline{\sigma_{\mathrm{clos}}}) = [\sigma_\baddr,\sigma_\eaddr] \tand \\
+    \code(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
+  \end{multlined}
+  \right\}
 \]
@@
 \[
-  H^\mathrm{code}_\sigma \; \ta \; s \; \sigma \; \hat{W} = 
+  H^\mathrm{code}_\sigma \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code \; s \; \sigma \; \hat{W} = 
   \begin{array}[t]{l}
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\sealed{\sigma,\retptrc(\rx,\lin,\baddr,\eaddr,\aaddr'+\calllen - 1)},\\((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\rx,\lin,\baddr,\eaddr,\aaddr'+\calllen - 1),\\((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
       \\
       \begin{array}{l}
-        \decInstr{\code([a',a' + \calllen]-1)} = \overline{\scall{o}{r_1}{r_2}} \tand \\
+        \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{o}{r_1}{r_2}} \tand \\
         \aaddr = \aaddr' + \retoffset \tand \\
-        \sigma = \sigma_\baddr + o \tand\\
+        \sigma = \sigma_\baddr + o \in \overline{\sigma_{\mathrm{ret}}}\tand\\
         \lbrack \baddr, \eaddr \rbrack = \dom(\code) \subseteq \ta
       \end{array}
-
     \end{array}
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\sealed{\sigma,\retptrd(\baddr,\eaddr)},\\((\rx,\linear),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
       \\
       \begin{array}{l}
-        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, ((\rx,\lin),\baddr',\eaddr',\aaddr')) \tand \\
+        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, ((\rx,\lin),\baddr',\eaddr',\aaddr'+\calllen)) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
-        \quad \decInstr{\code([a',a' + \calllen]-1)} = \overline{\scall{o}{r_1}{r_2}} \tand \\
-        \quad \aaddr = \aaddr' + \calllen \tand \\
-        \quad \sigma = \sigma_\baddr + o \tand\\
+        \quad \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{o}{r_1}{r_2}} \tand \\
+        \quad \sigma = \sigma_\baddr + o \in \overline{\sigma_{\mathrm{ret}}}\tand\\
+        \quad a = b-1\tand \\
         \quad \lbrack \baddr', \eaddr' \rbrack = \dom(\code) \subseteq \ta
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}
-      \sealed{\sigma,\vsc}, \vsc' \endarray\right)} \middle| \right. \\
-      \\
-      \begin{array}{l}
-        \sigma \in [\sigma_\baddr,\sigma_\eaddr] \tand\\
-        (\dom(\code) \cup \ta \neq \emptyset \tand \npair{\vsc,\vsc'} \in \lrv \; W) \tor \\
-        \text{not a call}
+\left. \npair{(\vsc, \vsc' )} \middle| \right. \\
+      \begin{array}{l}
+        \sigma \in \overline{\sigma_{\mathrm{clos}}} \tand\\
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{\vsc,\vsc'} \in \lrv \; W) \tor\\
+        (\dom(\code) \subseteq \ta)
       \end{array}

### 2018-03-20 fully-abstract-comp/technical-report.tex
Fix code region definition.

@@
 \[
-  \codereg{\ta,\overline{\sigma_{\mathrm{ret}}},\overline{\sigma_{\mathrm{clos}}},\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code,H^\mathrm{code}_\sigma \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code)
-\]
-
-\[
-  H^\mathrm{code} \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code \; s \; \hat{W} = \left\{
+  \codereg{\ta,\sigma_\baddr,\nrcalls,\nrclosures,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigma_\baddr \; \nrcalls \nrclosures \; \code)
+\]
+
+\[
+  H^\mathrm{code} \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code \; s \; \hat{W} = \left\{
     \npair{(\code,\code)} \mid
   \begin{multlined}
-    \dom(\code) = [\aaddr,\baddr] \tand (\overline{\sigma_{\mathrm{ret}}} \cup \overline{\sigma_{\mathrm{clos}}}) = [\sigma_\baddr,\sigma_\eaddr] \tand \\
+    \dom(\code) = [\aaddr,\baddr] \tand \sigma_\eaddr = \sigma_\baddr + \nrcalls+\nrclosures - 1 \tand \\
     \code(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
@@
 \lau{19-03-2018: Add the link capability to the above? code in ta?}
-\[
-  H^\mathrm{code}_\sigma \; \ta \; \overline{\sigma_{\mathrm{ret}}} \; \overline{\sigma_{\mathrm{clos}}} \; \code \; s \; \sigma \; \hat{W} = 
+\begin{multline*}
+  H^\mathrm{code}_\sigma \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code \; s \; \sigma \; \hat{W} = \\
   \begin{array}[t]{l}
@@
         \aaddr = \aaddr' + \retoffset \tand \\
-        \sigma = \sigma_\baddr + o \in \overline{\sigma_{\mathrm{ret}}}\tand\\
+        \sigma = \sigma_\baddr + o \tand o < \nrcalls\\
         \lbrack \baddr, \eaddr \rbrack = \dom(\code) \subseteq \ta
@@
         \quad \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{o}{r_1}{r_2}} \tand \\
-        \quad \sigma = \sigma_\baddr + o \in \overline{\sigma_{\mathrm{ret}}}\tand\\
+        \quad \sigma = \sigma_\baddr + o \tand o < \nrcalls \tand\\
         \quad a = b-1\tand \\
@@
       \begin{array}{l}
-        \sigma \in \overline{\sigma_{\mathrm{clos}}} \tand\\
+        \sigma = \sigma_\baddr + o \tand \nrcalls \leq o < \nrcalls + \nrclosures \tand\\
         (\dom(\code) \mathrel{\#} \ta \tand \npair{\vsc,\vsc'} \in \lrv \; W) \tor\\
@@
   \end{array}
-\]
+\end{multline*}
 

### 2018-03-20 fully-abstract-comp/technical-report.tex
code region change
formatting
First stab at a notion of components

@@
 \end{figure}
+
+\subsection{Components}
+\label{sec:components}
+
+We define a component as follows:
+\begin{align*}
+  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigma_\baddr,\nrcalls,\nrclosures)
+\end{align*}
+
+We say that a component is valid ($\vdash \var{comp}$), iff:
+\begin{mathpar}
+  \inference{}{
+    A_{\mathrm{code}},A_{\mathrm{data}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} z
+  }
+  \and
+  \inference{
+    c \in \overline{c_{\mathrm{import}}}
+  }{
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} c
+  }
+  \and
+  \inference{
+    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} & \perm \sqsubseteq \rx
+  }{
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
+  }
+  \and
+  \inference{
+    [\baddr,\eaddr] \subseteq A_{\mathrm{data}} & \permbnf \sqsubseteq \rw&
+    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}}
+  }{
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+  }
+  \and
+  \inference{
+    \dom(\ms_{\mathrm{code}}) = [\baddr_{\mathrm{code}},\eaddr_{\mathrm{code}}] &
+    \ms_{\mathrm{code}}(\baddr_{\mathrm{code}}) = \seal{\sigma_\baddr,\sigma_\baddr+\nrcalls+\nrclosures-1,\sigma_\baddr}\\
+    \forall a \in [\baddr_{\mathrm{code}}+1,\eaddr_{\mathrm{code}}]\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),\emptyset,\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}}\ms_{\mathrm{code}}(a)\\
+    \exists A_\mathrm{own} : \dom(\ms_{\mathrm{data}}) \rightarrow \powerset{\dom(\ms_{\mathrm{data}})} & \dom(\ms_{\mathrm{data}}) = \biguplus_{a \in \dom(\ms_{\mathrm{data}})} A_{\mathrm{own}}(a) \\
+    \forall a \in \dom(\ms_{\mathrm{data}})\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}}\ms_{\mathrm{data}}(a)\\
+  }{
+    \vdash (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigma_\baddr,\nrcalls,\nrclosures)
+  }
+\end{mathpar}
 
@@
         \sigma = \sigma_\baddr + o \tand \nrcalls \leq o < \nrcalls + \nrclosures \tand\\
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{\vsc,\vsc'} \in \lrv \; W) \tor\\
-        (\dom(\code) \subseteq \ta)
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; W) \tor\\
+        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; W)
       \end{array}

### 2018-03-20 fully-abstract-comp/technical-report.tex
changed cseal to use use source register as target
Valid use of offsets in a component.

@@
   \inference{}{
-    A_{\mathrm{code}},A_{\mathrm{data}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} z
+    A_{\mathrm{code}},A_{\mathrm{data}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} z
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} c
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} c
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+  }
+  \and
+  \inference{
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} \vsc \\
+    \sigma = \sigma_b + o & \nrcalls \leq o < \nrcalls + \nrclosures
+  }{
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
   }
@@
     \ms_{\mathrm{code}}(\baddr_{\mathrm{code}}) = \seal{\sigma_\baddr,\sigma_\baddr+\nrcalls+\nrclosures-1,\sigma_\baddr}\\
-    \forall a \in [\baddr_{\mathrm{code}}+1,\eaddr_{\mathrm{code}}]\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),\emptyset,\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}}\ms_{\mathrm{code}}(a)\\
+    \forall a \in [\baddr_{\mathrm{code}}+1,\eaddr_{\mathrm{code}}]\ldotp \ms_{\mathrm{code}}(a) \in \ints\\
+    \exists \var{offset\_use\_site} \in [0\cdots \nrcalls-1] \fun \dom(\ms_{\mathrm{code}}) \\
+    \forall a \in [\baddr_{\mathrm{code}}+1,\eaddr_{\mathrm{code}}-\calllen+1]\ldotp \ms_{\mathrm{code}}([a\cdots a+\calllen-1]) = \scall[0..\calllen-1]{o}{r_1}{r_2} \Rightarrow \var{offset\_use\_site}(o) =a\\
     \exists A_\mathrm{own} : \dom(\ms_{\mathrm{data}}) \rightarrow \powerset{\dom(\ms_{\mathrm{data}})} & \dom(\ms_{\mathrm{data}}) = \biguplus_{a \in \dom(\ms_{\mathrm{data}})} A_{\mathrm{own}}(a) \\
-    \forall a \in \dom(\ms_{\mathrm{data}})\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}} \vdash_{\mathrm{comp-value}}\ms_{\mathrm{data}}(a)\\
+    \forall a \in \dom(\ms_{\mathrm{data}})\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}}\ms_{\mathrm{data}}(a)\\
   }{
@@
 
-\[
-  H^\mathrm{code} \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code \; s \; \hat{W} = \left\{
-    \npair{(\code,\code)} \mid
+\begin{multline*}
+  H^\mathrm{code} \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code \; s \; \hat{W} = \\\left\{
+    \npair{(\code,\code)} \middle|
   \begin{multlined}
@@
   \right\}
-\]
+\end{multline*}
+
 \lau{19-03-2018: Add the link capability to the above? code in ta?}

### 2018-03-20 fully-abstract-comp/technical-report.tex
updated return seals use condition to take Dominique todo into account

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
       & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
-  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}}{\trg{r}} \mid \\ 
+  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}} \mid \\ 
       & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
@@
     \text{// seal the used stack frame as the data part of the return pointer pair.}\\
-    \tcseal{\rretd}{\rretd}{\rtmp{1}}\\
+    \tcseal{\rretd}{\rtmp{1}}\\
     \text{// obtain the code part of the return pointer pair and seal it too.}\\
@@
     \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
-    \tcseal{\rretc}{\rretc}{\rtmp{1}}\\
+    \tcseal{\rretc}{\rtmp{1}}\\
     \text{// now clear temporary register and jump to the adversary.}\\
@@
 \begin{align*}
-  \sem{\tcseal{r_1}{r_2}{r_3}}(\Phi) = &
+  \sem{\tcseal{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
@@
                                     \begin{array}{rl}
-                                      \Phi&\updReg{r_2}{w_2}\\
-                                          &\updReg{r_1}{\vsc}
+                                      \Phi&\updReg{r_1}{\vsc}
                                     \end{array}\right)
@@
                                     \begin{array}[t]{l}
-                                      \Phi(r_2) \in \SealableCaps \tand \\
-                                      \Phi(r_3) = \seal{\sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
-                                      \withinBounds{\Phi(r_3)} \tand \\
+                                      \Phi(r_1) \in \SealableCaps \tand \\
+                                      \Phi(r_2) = \seal{\sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
                                       \vsc = \sealed{\sigma,\Phi(r_2)} \tand\\
-                                      w_2 = \linCons{\Phi(r_2)}                                     
-                                    \end{array}
+                                      w_2 = \linCons{\Phi(r_2)}                                                           \end{array}
                                     \\
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_2)
-  = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\aaddr \in [\baddr,\eaddr]$, $\aaddr
-  \in \ta$ and $\Phi.\mem(\aaddr) = \seal{\sigma_b,\sigma_e}$, then
-  $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$.
-  \dominique{TODO: this should be allowed to be able to construct closures. Not allowed only for return pointer seals.}
+  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$ and $\sigma_\baddr \leq \sigma \leq \sigma_\baddr + \nrcalls$ or $\sigma_\baddr + \nrcalls < \sigma \leq \sigma_\baddr + \nrcalls + \nrclosures+1$.
 \end{definition}

### 2018-03-20 fully-abstract-comp/technical-report.tex
Finish component validity judgement.

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
 
-We say that a component is valid ($\vdash \var{comp}$), iff:
+We define inductively when a component is valid ($\vdash \var{comp}$) by the below inference rules:
 \begin{mathpar}
@@
     \forall a \in \dom(\ms_{\mathrm{data}})\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}}\ms_{\mathrm{data}}(a)\\
+    \overline{\dom(\ms_{\mathrm{code}}), \dom(\ms_{\mathrm{data}}), \emptyset, \overline{c_{\mathrm{import}}}, \sigma_\baddr, \nrcalls, \nrclosures \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}
   }{

### 2018-03-20 fully-abstract-comp/technical-report.tex
off by 1 and 2 errors

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$ and $\sigma_\baddr \leq \sigma \leq \sigma_\baddr + \nrcalls$ or $\sigma_\baddr + \nrcalls < \sigma \leq \sigma_\baddr + \nrcalls + \nrclosures+1$.
+  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$ and $\sigma_\baddr \leq \sigma < \sigma_\baddr + \nrcalls$ or $\sigma_\baddr + \nrcalls \leq \sigma < \sigma_\baddr + \nrcalls + \nrclosures$.
 \end{definition}

### 2018-03-21 fully-abstract-comp/technical-report.tex
sanity conditions

@@
   \]
+  and
+  \[
+    r_1' \neq \rtmp{1}
+  \]
   % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
@@
 \begin{definition}[Use return seals only for call]
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\compbaseseal$, $\nrcalls$, and $\nrclosures$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$ and $\sigma_\baddr \leq \sigma < \sigma_\baddr + \nrcalls$ or $\sigma_\baddr + \nrcalls \leq \sigma < \sigma_\baddr + \nrcalls + \nrclosures$.
+  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$ and $\compbaseseal \leq \sigma < \compbaseseal + \nrcalls$ or $\compbaseseal + \nrcalls \leq \sigma < \compbaseseal + \nrcalls + \nrclosures$.
 \end{definition}
@@
 \begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\compbaseseal$, $\nrcalls$, and $\nrclosures$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
   \]
-  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and
-  $\src{\Phi}.\reg(r_i) = \seal{\sigma_b,\sigma_e}$ and
-  \dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
-  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
-  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}
-  \dominique{TODO: Also not allowed: sealing such a value with a seal that you do not own.}
+  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
+  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
+  \[
+    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow [\compbaseseal,\compbaseseal + \nrcalls + \nrclosures - 1] \# [\sigma_b,\sigma_e]
+  \]
+and
+%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
+  \[
+    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
+  \]
+  and
+    \begin{multline*}
+      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in [\compbaseseal,\compbaseseal + \nrcalls + \nrclosures - 1] \Rightarrow 
+      \\  \left(  
+      \begin{array}{r l}
+        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow [\compbaseseal,\compbaseseal + \nrcalls + \nrclosures - 1] \# [\sigma_b,\sigma_e]) \wedge\\
+        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
+      \end{array}
+      \right)
+  \end{multline*}
 \end{definition}
+
+%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
+\begin{definition}[Never store link seals and code capability]
+  A configuration $\src{\Phi}$ is said to not store seals and code capability in $\ta$ with component parameters $\compbaseseal$, $\nrcalls$, and $\nrclosures$ for $n$ steps iff\\
+  for all $i \leq n$, if
+  \[
+    \src{\Phi} \nstep[i] \src{\Phi'}
+  \]
+  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
+\[
+  \Phi'(r_2) = \seal(\sigma_\baddr,\sigma_\eaddr,\_) \Rightarrow [\compbaseseal, \compbaseseal + \nrcalls  + \nrclosures - 1] \# [\sigma_\baddr,\sigma_\eaddr]
+\]
+and
+\[
+  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
+\]
+\end{definition}
+
+%   \dominique{TODO: Also not allowed: sealing such a value with a seal that you do not own.}
+%   \lau{is this covered by definition 14 (Use return seals only for call)? }
+% \begin{definition}[Never seal link seals and code capabilities with unknown seal]  
+% \end{definition}
 

### 2018-03-21 fully-abstract-comp/technical-report.tex
more sanity condition related changes

@@
   \label{def:points-to-instr}
-  A configuration $(\ms,\reg)$ {\sourcecolor or $(\ms,\reg,\_,\_)$} is said to point instruction $i$ in $A$ iff
-  \[
-    \reg(\pcreg) = ((\_,\_),\_,\_,\aaddr) \tand \aaddr \in A
+  A configuration $(\ms,\reg)$ {\sourcecolor or $\Phi = (\ms,\reg,\_,\_)$} is said to point to non-$\scall{}{}{}$ instruction $i$ in $A$ iff
+  \[
+    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand \aaddr \in A
   \]
@@
 
+\begin{definition}
+  \label{def:points-to-call}
+  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{o}{r_1}{r_2}$ in $A$ iff
+  \[
+    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand [\aaddr,\aaddr+\calllen-1] \subseteq A
+  \]
+and
+\[
+  \aaddr+\calllen - 1 \leq \eaddr
+\]
+and
+\[
+  \callCond{\Phi,r_1,r_2,\aaddr}
+\]
+\end{definition}
 We want to make a conditional full-abstraction statement.
 The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do.
-(2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation
+(2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation.
 \begin{definition}
   \label{def:check-stack-addr-before-call}
-  A configuration $\src{\Phi}$ is said to have reasonable calls in $\ta$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
   for all $i \leq n$ if
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\scall{o}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2}$
+  and $\src{\Phi'}$ points to $\src{\scall{o}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
   % and there exists no $j<i$ s.t.

### 2018-03-21 fully-abstract-comp/technical-report.tex
Modify convention about seal locations
In more detail:
* call now takes an offset saying where the seal can be found relative to the pc
* components can now put seals everywhere they want, not just at the start of their
  code

Reason: makes linking easier, simpler overall.

--- 2018-03-20 fully-abstract-comp/technical-report.tex
@@
     o & \in & \nats \\
-    \src{i} & \defbnf &  \trg{i} \mid \scall{o}{r}{r}
-  \end{array}
-\]
-There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{o}{}{}$ instruction (the $o$ refers to the seal it uses, namely the offset in the seals made available by linking).
-
-The source machine allows for variable length instructions which we utilise for $\scall{o}{}{}$.
-In other words, when $\scall{o}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction (and be within the range of the pc capability).
+    \src{i} & \defbnf &  \trg{i} \mid \scall{\offpc,\offsigma}{r}{r}
+  \end{array}
+\]
+There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{\offpc,\offsigma}{}{}$ instruction (the $o$ refers to the seal it uses, namely the offset in the seals made available by linking).
+
+The source machine allows for variable length instructions which we utilise for $\scall{\offpc,\offsigma}{}{}$.
+In other words, when $\scall{\offpc,\offsigma}{}{}$ is decoded, a series of addresses in memory must contain what corresponds to the call instruction (and be within the range of the pc capability).
 This is described in more detail when we present the decoding function.
@@
 \paragraph{Step relation}
-The instruction $\scall{o}{}{}$ has length $\calllen$ which means that it does not fit in one memory address.
-In fact, $\scall{o}{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
-We therefore introduce $\scall[0]{o}{r_1}{r_2},\dots,\scall[\calllen-1]{o}{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{o}{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
-We define the following condition that indicates that $\aaddr$ is the first address of a $\scall{o}{r_1}{r_2}$ instruction in the configuration $\Phi$:
+The instruction $\scall{\offpc,\offsigma}{}{}$ has length $\calllen$ which means that it does not fit in one memory address.
+In fact, $\scall{\offpc,\offsigma}{}{}$ should be seen as a different way of interpreting a series of instruction rather than an instruction on its own.
+We therefore introduce $\scall[0]{\offpc,\offsigma}{r_1}{r_2},\dots,\scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}$ as aliases for the instructions that constitute $\scall{\offpc,\offsigma}{r_1}{r_2}$ (see Paragraph~\ref{par:call-impl} for details).
+We define the following condition that indicates that $\aaddr$ is the first address of a $\scall{\offpc,\offsigma}{r_1}{r_2}$ instruction in the configuration $\Phi$:
 \[
@@
       \begin{array}{l}
-        \Phi.\mem(\aaddr) = \scall[0]{o}{r_1}{r_2} \tand\\
+        \Phi.\mem(\aaddr) = \scall[0]{\offpc,\offsigma}{r_1}{r_2} \tand\\
         \vdots \\
-        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen-1]{o}{r_1}{r_2}
+        \Phi.\mem(\addr+\calllen-1) = \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}
       \end{array}
@@
 \begin{align*}
-  \src{\Phi} & \; \src{\step \sem{\scall{o}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
+  \src{\Phi} & \; \src{\step \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                   \begin{array}[t]{l}
@@
 \label{par:call-impl}
-This paragraph contains the implementation of $\scall{o}{}{}$.
-That is each of the instructions in the implementation corresponds to $\scall[1]{o}{r_1}{r_2} \dots \scall[\calllen-1]{o}{r_1}{r_2}$, respectively.
-
-%\lau{16-10-2017: Fill this paragraph with $\scall{o}{}{}$ implementation, i.e., the instructions $\scall[1]{o}{r_1}{r_2} \dots \scall[\calllen]{o}{r_1}{r_2}$ is equal to.
-Add convenience notation like $\update{\mem.a}{\scall{o}{r_1}{r_2}} = \dots$}
-\lau{16-10-2017: Implementation of $\scall{o}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
+This paragraph contains the implementation of $\scall{\offpc,\offsigma}{}{}$.
+That is each of the instructions in the implementation corresponds to $\scall[1]{\offpc,\offsigma}{r_1}{r_2} \dots \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}$, respectively.
+
+%\lau{16-10-2017: Fill this paragraph with $\scall{\offpc,\offsigma}{}{}$ implementation, i.e., the instructions $\scall[1]{\offpc,\offsigma}{r_1}{r_2} \dots \scall[\calllen]{\offpc,\offsigma}{r_1}{r_2}$ is equal to.
+Add convenience notation like $\update{\mem.a}{\scall{\offpc,\offsigma}{r_1}{r_2}} = \dots$}
+\lau{16-10-2017: Implementation of $\scall{\offpc,\offsigma}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
 \[
@@
     \tmove{\rtmp{1}}{\pcreg}\\
-    \tsetatob{\rtmp{1}}\\
+    \tcca{\rtmp{1}}{\offpc} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code}\\
     \tload{\rtmp{1}}{\rtmp{1}}\\
-    \tcca{\rtmp{1}}{o}\text{ // $o$ is the offset this call uses.}\\
+    \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
     \text{// seal the used stack frame as the data part of the return pointer pair.}\\
@@
 \dominique{Return code could be simplified if we had jz in addition to jnz.}
-\dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
-  I understand this is for ensuring that our private stack frame is not empty, which is important.
-  But why don't we just push a constant then?
-  Also: perhaps we could only do it if the current private stack frame is empty to begin with?
-}
-\dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
+% \dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
+%   I understand this is for ensuring that our private stack frame is not empty, which is important.
+%   But why don't we just push a constant then?
+%   Also: perhaps we could only do it if the current private stack frame is empty to begin with?
+% }
+% \dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
 The call code does the following:
@@
 
-For convenience, we will add the convention that the memory update $\update{\mem.a}{\scall{o}{r_1}{r_2}}$ corresponds to
-\[
-  \begin{array}{l}
-    \update{\mem.a}{\scall[0]{o}{r_1}{r_2}}\\
-    \update{\mem.a+1}{\scall[1]{o}{r_1}{r_2}}\\
+For convenience, we will add the convention that the memory update $\update{\mem.a}{\scall{\offpc,\offsigma}{r_1}{r_2}}$ corresponds to
+\[
+  \begin{array}{l}
+    \update{\mem.a}{\scall[0]{\offpc,\offsigma}{r_1}{r_2}}\\
+    \update{\mem.a+1}{\scall[1]{\offpc,\offsigma}{r_1}{r_2}}\\
     \dots\\
-    \update{\mem.a+\calllen-1}{\scall[\calllen-1]{o}{r_1}{r_2}}\\
+    \update{\mem.a+\calllen-1}{\scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}\\
   \end{array}
@@
 \begin{align*}
-  \sem{\scall{o}{\src{r_1}}{\src{r_2}}}(\Phi) = & 
+  \sem{\scall{\offpc,\offsigma}{\src{r_1}}{\src{r_2}}}(\Phi) = & 
                                                \begin{cases}
@@
                                                    \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
-                                                   \mem(\baddr) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
-                                                   \sigma_? = \sigma_\aaddr + o \tand\\
+                                                   \mem(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
+                                                   \sigma_? = \sigma_\aaddr + \offsigma \tand\\
                                                    \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
@@
 
-\subsection{Program layout and linking}
+\subsection{Component layout and linking}
+We define a notion of
+
 A target-level program needs a number of things available in order to be able to execute securely. In the following, we list what it needs followed by a description of how they are made available.
@@
 \begin{align*}
-  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigma_\baddr,\nrcalls,\nrclosures)
+  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
 \end{align*}
@@
 \begin{mathpar}
-  \inference{}{
-    A_{\mathrm{code}},A_{\mathrm{data}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} z
+  \inference{
+    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} & [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)
+  }{
+    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+  }
+  \and
+  \inference{
+    \mscode(a) \in \ints\\
+    \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge \sigma_\baddr+\offsigma \in \sigrets)
+  }{
+    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+  }
+  \and
+  \inference{
+    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
+    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
+    \forall a \in \dom(\mscode) \ldotp 
+    d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+  }{
+    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode
+  }
+  \and
+  \inference{
+  }{
+    A_{\mathrm{code}},A_{\mathrm{data}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} c
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
   }
@@
   \inference{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} \vsc \\
-    \sigma = \sigma_b + o & \nrcalls \leq o < \nrcalls + \nrclosures
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
+    \sigma \in \sigcloss
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
   }
@@
   \inference{
-    \dom(\ms_{\mathrm{code}}) = [\baddr_{\mathrm{code}},\eaddr_{\mathrm{code}}] &
-    \ms_{\mathrm{code}}(\baddr_{\mathrm{code}}) = \seal{\sigma_\baddr,\sigma_\baddr+\nrcalls+\nrclosures-1,\sigma_\baddr}\\
-    \forall a \in [\baddr_{\mathrm{code}}+1,\eaddr_{\mathrm{code}}]\ldotp \ms_{\mathrm{code}}(a) \in \ints\\
-    \exists \var{offset\_use\_site} \in [0\cdots \nrcalls-1] \fun \dom(\ms_{\mathrm{code}}) \\
-    \forall a \in [\baddr_{\mathrm{code}}+1,\eaddr_{\mathrm{code}}-\calllen+1]\ldotp \ms_{\mathrm{code}}([a\cdots a+\calllen-1]) = \scall[0..\calllen-1]{o}{r_1}{r_2} \Rightarrow \var{offset\_use\_site}(o) =a\\
-    \exists A_\mathrm{own} : \dom(\ms_{\mathrm{data}}) \rightarrow \powerset{\dom(\ms_{\mathrm{data}})} & \dom(\ms_{\mathrm{data}}) = \biguplus_{a \in \dom(\ms_{\mathrm{data}})} A_{\mathrm{own}}(a) \\
-    \forall a \in \dom(\ms_{\mathrm{data}})\ldotp \dom(\ms_{\mathrm{code}}),\dom(\ms_{\mathrm{data}}),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}},\sigma_\baddr,\nrcalls,\nrclosures \vdash_{\mathrm{comp-value}}\ms_{\mathrm{data}}(a)\\
-    \overline{\dom(\ms_{\mathrm{code}}), \dom(\ms_{\mathrm{data}}), \emptyset, \overline{c_{\mathrm{import}}}, \sigma_\baddr, \nrcalls, \nrclosures \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}
+    \dom(\mscode) \mathrel{\#} \dom(\msdata)\\
+    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
+    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
+    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),\dom(\msdata),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash \msdata(a)\\
+    \overline{\dom(\mscode), \dom(\msdata), \emptyset, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
   }{
-    \vdash (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigma_\baddr,\nrcalls,\nrclosures)
+    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+  }
+\end{mathpar}
+
+\subsection{Linking}
+\label{sec:linking}
+
+TODO:
+\begin{mathpar}
+  \inference{
+  }{
+    \var{comp}_1 \bowtie \var{comp}_2 = \var{comp}_3
   }
@@
 \[
-  \comp{\scall{o}{r_1}{r_2}} = 
+  \comp{\scall{\offpc,\offsigma}{r_1}{r_2}} = 
   \begin{array}[t]{l}
@@
 \[
-  \codereg{\ta,\sigma_\baddr,\nrcalls,\nrclosures,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigma_\baddr \; \nrcalls \nrclosures \; \code)
-\]
-
-\begin{multline*}
-  H^\mathrm{code} \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code \; s \; \hat{W} = \\\left\{
-    \npair{(\code,\code)} \middle|
-  \begin{multlined}
-    \dom(\code) = [\aaddr,\baddr] \tand \sigma_\eaddr = \sigma_\baddr + \nrcalls+\nrclosures - 1 \tand \\
-    \code(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}
-  \end{multlined}
+  \codereg{\ta,\sigrets,\sigcloss} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigrets \; \sigcloss \; \code)
+\]
+
+\begin{equation*}
+  H^\mathrm{code} \; \ta \; \sigrets \; \sigcloss \; \code \; s \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
+    \dom(\code) = [\aaddr,\baddr]
   \right\}
-\end{multline*}
+\end{equation*}
 
@@
 \begin{multline*}
-  H^\mathrm{code}_\sigma \; \ta \; \sigma_\baddr \; \nrcalls \; \nrclosures \; \code \; s \; \sigma \; \hat{W} = \\
+  H^\mathrm{code}_\sigma \; \ta \; \sigrets \; \sigcloss \; \code \; s \; \sigma \; \hat{W} = \\
   \begin{array}[t]{l}
@@
       \begin{array}{l}
-        \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{o}{r_1}{r_2}} \tand \\
+        \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
         \aaddr = \aaddr' + \retoffset \tand \\
-        \sigma = \sigma_\baddr + o \tand o < \nrcalls\\
-        \lbrack \baddr, \eaddr \rbrack = \dom(\code) \subseteq \ta
+        \code(a+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
+        \lbrack \baddr, \eaddr \rbrack \subseteq \dom(\code) \subseteq \ta
       \end{array}
@@
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
-        \quad \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{o}{r_1}{r_2}} \tand \\
-        \quad \sigma = \sigma_\baddr + o \tand o < \nrcalls \tand\\
+        \quad \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \quad \code(a+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
         \quad a = b-1\tand \\
-        \quad \lbrack \baddr', \eaddr' \rbrack = \dom(\code) \subseteq \ta
+        \quad \lbrack \baddr', \eaddr' \rbrack \subseteq \dom(\code) \subseteq \ta
       \end{array}
@@
       \begin{array}{l}
-        \sigma = \sigma_\baddr + o \tand \nrcalls \leq o < \nrcalls + \nrclosures \tand\\
+        \sigma \in \sigcloss \tand\\
         (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; W) \tor\\
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\scall{o}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2}$
+  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2}$
   % and there exists no $j<i$ s.t.
@@
   % \]
-  % where $\src{\Phi''}$ points to $\src{\scall{o}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
+  % where $\src{\Phi''}$ points to $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
   then
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{o}{r_1'}{r_2'}}$ and $\sigma_\baddr \leq \sigma < \sigma_\baddr + \nrcalls$ or $\sigma_\baddr + \nrcalls \leq \sigma < \sigma_\baddr + \nrcalls + \nrclosures$.
+  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ or $\sigma \in \sigcloss$.
 \end{definition}
@@
 That said, I would like to keep it around (maybe change it so it becomes an argument for variable length instructions).}
-This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{o}{}{}$ and $\sreturn$ instructions.
+This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{\offpc,\offsigma}{}{}$ and $\sreturn$ instructions.
 The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.
@@
 \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-$\scall{o}{r5}{0}$
+$\scall{\offpc,\offsigma}{r5}{0}$
 $\sjmp{r5}$
 \end{lstlisting}
-\footnote{At this time $\scall{o}{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
+\footnote{At this time $\scall{\offpc,\offsigma}{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
 We would not have enough instructions to set up a proper return pointer for the compiled return to use.
@@
 We do not offer primitive \sreturn{}, \spush{} and \spop{} instructions, even though there is nothing keeping us from doing that, either as an additional primitive instruction or as a macro on top of the primitive token values mentioned above.
-We do offer a primitive \scall{o}{}{} instruction, and it offers the only way to create a new stack frame.
+We do offer a primitive \scall{\offpc,\offsigma}{}{} instruction, and it offers the only way to create a new stack frame.
 It is the only instruction that does not have a target language counterpart.
-However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{o}{}{} compiles to.
+However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{\offpc,\offsigma}{}{} compiles to.
 We will discuss later that this means that if we back translate an adversary piece of code (in the full abstraction proof), then it will be back translated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
@@
 % Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
-% \lau{07-09-2017: One would use $\scall{o}{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
+% \lau{07-09-2017: One would use $\scall{\offpc,\offsigma}{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
 The return would happen just by using $\sxjmp{}{}$ with the $\retptrd$ and $\retptrc$.
@@
 In other words, we need something to recognise when a return happens.}
-% \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{o}{}{}$ instruction.
+% \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{\offpc,\offsigma}{}{}$ instruction.
 The real implementation of the $\retptrc$ will simply point to a piece of clean up code that splices together the old/private stack pointer with the one we got from the untrusted party.
-This may, however, give some issues when we want to reason about it because there will be a disconnect between where the clean up job is done in the operational semantics of the source language (namely as part of the $\sxjmp{}{}$), but in a translated program, the code that does the clean up origins with the $\scall{o}{}{}$ instruction.
+This may, however, give some issues when we want to reason about it because there will be a disconnect between where the clean up job is done in the operational semantics of the source language (namely as part of the $\sxjmp{}{}$), but in a translated program, the code that does the clean up origins with the $\scall{\offpc,\offsigma}{}{}$ instruction.
 }
@@
 The reason is that we only want to prove a result about programs that respect the calling convention, i.e.\ that do not do anything stupid.
-For example, in the old setup with global/local capabilities, when trusted code performs an \scall{o}{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
+For example, in the old setup with global/local capabilities, when trusted code performs an \scall{\offpc,\offsigma}{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
 Such sanity conditions are necessary (we cannot guarantee anything for code that does such stupid things) but we do not want to spend computation time on checking them.

### 2018-03-21 fully-abstract-comp/technical-report.tex
comment

@@
   \]
-  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $A$ for some $\src{r_1}$ and $\src{r_2}$
+  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
   % and there exists no $j<i$ s.t.
@@
 \begin{definition}[Use return seals only for call]
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ or $\sigma \in \sigcloss$.
+  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either ($\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$) or $\sigma \in \sigcloss$.
 \end{definition}
@@
 \begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
   \]
-  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and
-  $\src{\Phi}.\reg(r_i) = \seal{\sigma_b,\sigma_e}$ and
-  \dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
-  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
-  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}
-  \dominique{TODO: Also not allowed: sealing such a value with a seal that you do not own.}
+  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
+  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
+  \[
+    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow (\sigrets \cup \sigcloss) \mathrel{\#} [\sigma_b,\sigma_e]
+  \]
+and
+%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
+  \[
+    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
+  \]
+  and
+    \begin{multline*}
+      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in (\sigrets\cup\sigcloss) \Rightarrow 
+      \\  \left(  
+      \begin{array}{r l}
+        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow (\sigrets\cup\sigcloss) \# [\sigma_b,\sigma_e]) \wedge\\
+        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
+      \end{array}
+      \right)
+  \end{multline*}
 \end{definition}
+
+%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
+\begin{definition}[Never store link seals and code capability]
+  A configuration $\src{\Phi}$ is said to not store seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  for all $i \leq n$, if
+  \[
+    \src{\Phi} \nstep[i] \src{\Phi'}
+  \]
+  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
+\[
+  \Phi'(r_2) = \seal(\sigma_\baddr,\sigma_\eaddr,\_) \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
+\]
+and
+\[
+  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
+\]
+\end{definition}
+
+%   \dominique{TODO: Also not allowed: sealing such a value with a seal that you do not own.}
+%   \lau{is this covered by definition 14 (Use return seals only for call)? }
+% \begin{definition}[Never seal link seals and code capabilities with unknown seal]  
+% \end{definition}
 

### 2018-03-21 fully-abstract-comp/technical-report.tex
Remove old value relation.
fix seal offset bug in implementation of call

@@
     \tmove{\rtmp{1}}{\pcreg}\\
-    \tcca{\rtmp{1}}{\offpc} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code}\\
+    \tcca{\rtmp{1}}{(\offpc - 5)} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code.
+\footnote{-5 is the offset from the previous instruction where the $\pcreg$ was moved to $\rtmp{1}$ to the first instruction of the call implementation.}}\\
     \tload{\rtmp{1}}{\rtmp{1}}\\
+
     \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
@@
   \]
+  \lau{maybe the last bit should have its own definition.}
   % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.

### 2018-03-21 fully-abstract-comp/technical-report.tex
updated definition of 'reasonable' component used in full-abstraction premise

@@
 
-% \[
-%   \lrv(H_\sigma,W) =
-%   \begin{array}[t]{l}
-%     \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
-% %
-%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr)},\\
-%                          \sealed{\sigma,((\perm,\lin),\baddr,\eaddr,\aaddr + ?)}\endarray\right)} \;\middle| \;
-%     \begin{array}{l}
-%       (\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C}) \in H_\sigma \; \sigma \wedge \\
-%       \npair{([\retptrc(\sigma,\perm,\lin,\baddr,\eaddr,\aaddr),\sealed{\sigma,\vsc_C})]} \in \lrp(W)
-%     \end{array}
-%     \right\}\cup\\
-% %
-%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,(\baddr,\eaddr))},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)} \;\middle| \;
-%     \begin{array}{l}
-%       \npair{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W) \wedge \\
-%       \forall n' \leq n \ldotp \forall W' \privft W , W_C,W_\lrm,W_\lrr\ldotp \\
-%       \quad \npair[n']{\left[\left(\arraycolsep=0pt\array{l}\src{\retptrd(\sigma,\baddr,\eaddr)},\\\sealed{\sigma,((\rw,\linear),\baddr,\eaddr,\baddr)}\endarray\right)\right]} \in \lrp(W') \wedge \\
-%       \quad \npair[n']{\left(\arraycolsep=0pt\array{l}\retptrc(\perm_C,\lin_C,\baddr_C,\eaddr_C,\aaddr_C),\\
-%                                                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C)\endarray\right)} \in \lrv(W_C) \wedge\\
-%       \quad \memSat[n']{\stpair[.]{\ms_S,(\ms_f,\opc)::\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
-%       \quad \opc = ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C) \wedge \\
-%       \quad \npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rtmp{1},\rtmp{2}\})(W_\lrr) \wedge \\
-%       \quad \src{\reg_S}(r_\stk) = \stkptr{\rw,\stkb,\baddr-1,\_} \wedge \\
-%       \quad \reg_T(r_\stk) = ((\rw,\linear),\stkb,\baddr-1,\_) \wedge \\
-%       \quad W'(r_\var{free}) = \stdreg{[\stkb,\baddr-1]}{\spa} \wedge \\
-%       \quad W'(r_\var{priv}) = \stareg[\src{\ms_f},\ms_f]{\spao} \wedge \\
-%       \quad \src{\ms_f} = \ms_f \wedge \dom(\ms_f) = [\baddr,\eaddr] \wedge\\
-%       \quad W_\var{new} = (W' \oplus W_\lrr \oplus W_\lrm \oplus W_C)[\revoked,\revoked,\stdreg{[\stkb,\eaddr]}{\spao}]\\
-%       \qquad \Rightarrow \\
-%       \qquad \quad \npair[n']{\left(\arraycolsep=0pt\array{l}
-%                                      \left(\arraycolsep=0pt\array{l}\ms_S,\reg_S \left[
-%       \array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\\stkptr{\rw,\stkb,\eaddr,\baddr},((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),0,0,0,0\endarray \right]\\
-%                                      ,\stk,\ms_\stk \uplus \ms_f
-%                                      \endarray \right),\\
-%                                              \left(\ms_T,\reg_S\left[\array{l}\reg.\rstk,\pcreg,\rdata,\rtmp{1},\rtmp{2},r_2\mapsto\\((\rw,\linear),\stkb,\eaddr,\baddr),((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C+?),0,0,0,0\endarray\right]\right)
-%                                     \endarray\right)} \in \lro
-%     \end{array} 
-%     \right\}\cup \\
-% %
-%     \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
-%     \begin{array}{l}
-%       \stpair{\vsc}{\vsc} \in H_\sigma \; \sigma \wedge \\
-%       \forall W' \pubft W, n' \leq n \ldotp \\
-%       \quad \forall \stpair{\vsc'}{\vsc'} \in H_\sigma \; \sigma \ldotp \\
-%       \qquad \npair[n']{(\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T')} \in \lre(W') \wedge \\
-%       \qquad \npair[n']{(\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T)} \in \lre(W')
-%     \end{array}
-%     \right\}\cup\\
-% %
-%     % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
-%     % \right\} \cup \\
-% %
-%     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
-%     \begin{array}{r l l}
-%       \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\linear,W}\\
-%       \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\linear,W}
-%     \end{array}
-%     \right\} \cup \\
-% %
-%     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
-%     \begin{array}{r l l }
-%       \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}\\
-%       \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W}\\
-%       \perm = \rwx &\Rightarrow& \npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}\\
-%       \perm = \rx &\Rightarrow& \npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
-%     \end{array}
-%     \right\}
-%   \end{array}
-% \]
-
-% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
-It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
 \[
@@
 \begin{definition}[Use return seals only for call]
+  \label{def:use-return-seals-call}
   A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
@@
 \begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  \label{def:handle-trusted-seals-properly}
+  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
 \begin{definition}[Never store link seals and code capability]
-  A configuration $\src{\Phi}$ is said to not store seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  \label{def:never-store-seal-code-cap}
+  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
 \[
-  \Phi'(r_2) = \seal(\sigma_\baddr,\sigma_\eaddr,\_) \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
+  \Phi'(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
 \]
@@
 \begin{definition}
-  A program $\program$ with memory segment $\ms_\program$ and $\dom(\ms_\program) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
-  \begin{itemize}
-  \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
-  \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
-  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
+  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  \begin{itemize}
+  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\plug{\context}{\program}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\plug{\context}{\program}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
   \end{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex
Define linking, programs, contexts.

@@
     \tmove{\rtmp{1}}{\pcreg}\\
-    \tcca{\rtmp{1}}{(\offpc - 5)} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code.
-\footnote{-5 is the offset from the previous instruction where the $\pcreg$ was moved to $\rtmp{1}$ to the first instruction of the call implementation.}}\\
+    \tcca{\rtmp{1}}{\offpc} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code}\\
     \tload{\rtmp{1}}{\rtmp{1}}\\
-
     \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
@@
 
-TODO:
 \begin{mathpar}
   \inference{
+    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
+    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
+    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
+    \mscode[3] = \mscode[1] \uplus \mscode[2] &
+    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
+    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}&
+    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}\\
+    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
+    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
   }{
-    \var{comp}_1 \bowtie \var{comp}_2 = \var{comp}_3
+     \var{comp}_3 = \var{comp}_1 \bowtie \var{comp}_2
   }
@@
 
-\subsection{Programs and contexts}
-\begin{definition}
-  \label{def:self-contained}
-  \begin{itemize}
-  \item We say that a memory segment $\ms$ and a list of words $\bar{w}$
-    are self-contained iff
-  \[
-    \forall w \in \bar{w} \ldotp w \in \Caps \lor w = \stkptr{\_,\_,\_,\_} \Rightarrow \range{w} \subseteq \dom(\ms)
-  \]
-  and
-  \begin{multline*}
-    \forall a \in \dom(\ms) \ldotp\\
-    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom(\ms)
-  \end{multline*}
-  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$ are self contained with slack $\ms_s$ iff
-  \[
-    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom(\ms \cup \ms_s)
-  \]
-  and
-  \begin{multline*}
-    \forall a \in \dom(\ms) \ldotp\\
-    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom(\ms \cup \ms_s)
-  \end{multline*}
-\end{itemize}
+\subsection{Programs, contexts, initial execution configuration}
+\label{sec:programs}
+
+\begin{definition}[Programs and Contexts]
+  We define a program as a $(\var{comp},c_{\mathrm{main}})$ component with an empty import list and $c_{\mathrm{main}}$ is one of the exported symbols.
+
+  Given a component $\var{comp}$, a context is a tuple $(\var{comp'},c_{\mathrm{main}})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'},c_{\mathrm{main}})$ is a program.
 \end{definition}
-
-\begin{definition}
-  \label{def:wxorx}
-  ~\\
-  A word $w$ is said to respect write-xor-execute iff
-  \begin{itemize}
-  \item $w \in \ints$
-  \item $w = \seal{\_,\_,\_}$
-  \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
-  \item $\sealed{\_,\vsc}$ and $\vsc$ respects write-xor-execute
-  \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
-  \item $\src{w = \retptrc(\_)}$
-  \item $\src{w = \retptrd(\_,\_)}$
-  \end{itemize}
-~\\
-  A register file $\reg$ is said to respect write-xor-execute with slack $[\bar{r}]$ iff \\
-  For all $r \in \RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
-~\\
-~\\
-  A memory segment $\ms$ is said to respect write-xor-execute iff\\
-  For all $\aaddr \in \dom(\ms)$ $\ms(\aaddr)$ respects write-xor-execute
-\end{definition}
-
-\begin{definition}
-  \label{def:no-stk-ptr}
-~\\
-  We say that a word $w$ covers no stack pointer iff\\
-  \begin{item}
-    \item $w \neq ((\perm,\linear),\stkb,\_,\_)$\\
-          $\src{w \neq \stkptr{\_,\_,\_,\_}}$
-    \item if $w = \sealed{\_,\vsc}$, then $\vsc$ covers no stack pointer.
-  \end{item}
-~\\
-  We say that a memory segment $\ms$ contains no stack pointer iff
-  \[
-    \forall \aaddr \in \dom(\ms)\ldotp \text{$\ms(\aaddr)$ covers no stack pointer}
-  \]
-~\\
-  We say that a register file $\reg$ contains no stack pointer with slack $[\bar{r}]$ iff
-  \[
-    \forall r \in \reg \setminus [\bar{r}] \ldotp \text{$\reg(r)$ covers no stack pointer}
-  \]
-\end{definition}
-
-\begin{definition}
-  \label{def:resp-linearity}
-~\\
-  We say that a memory segment $\ms$ and words $\bar{w}$ respects linearity iff\\
-  Let $\var{lc}$ be all the linear capabilities in $\ms$ and $\bar{w}$, i.e.,
-  \[
-    \var{lc} = \left\{c \middle|
-      \begin{array}{l}
-        \exists \aaddr \in \dom(\ms) \ldotp c = \ms(\aaddr) = ((\_,\linear),\_,\_,\_) \\
-        \exists \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) = \sealed{\_,c} \land c = ((\_,\linear),\_,\_,\_) \\
-        \src{\exists \aaddr \in \dom(\ms) \ldotp c = \ms(\aaddr) = \stkptr{\_,\_,\_,\_}} \\
-        \src{\exists \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) = \sealed{\_,c} \land c= \stkptr{\_,\_,\_,\_}} \\
-      \end{array}
-\right\}
-  \]
-\end{definition}
-
-\subsubsection{Programs}
-Here we define what constitutes a program.
-\lau{05-10-2017: It seems like both program and context takes care of linking atm.}
-\begin{definition}
-  \label{def:program}
-  \lau{27-09-2017: The ``somehow''s need to be fixed.}
-  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{\sigma,c_\code}$ and $\vsc_\data = \sealed{\sigma,c_\data}$, and two memory segments $\ms$ and $\ms_\link$ satisfying
-  \begin{itemize}
-  \item $c_\code = ((\rx,\lin_\code),\baddr_\code,\eaddr_\code,\aaddr_\code)$ where $\aaddr = \baddr + 2$ and $[\baddr_\code,\eaddr_\code]\subseteq \dom(\ms)$
-  \item $c_\data = ((\rw,\lin_\data),\baddr_\data,\eaddr_\data,\baddr_\data)$ where $[\baddr_\data,\eaddr_\data] \subseteq \dom(\ms)$ and $[\baddr_\code,\eaddr_\code] \cap [\baddr_\data,\eaddr_\data] = \emptyset$
-  \item $\ms$ and $[c_\code,c_\data]$ are self-contained with $\ms_\link$ as slack.
-  \item $\ms$ respects write-xor-execute.
-  \item $\ms(\baddr_\code) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ where these are somehow the sufficient amount of seals.
-  \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
-  \end{itemize}
-\end{definition}
-\lau{27-09-2017: At the moment, we do not have the proper support to let the code capability be linear (in the sense that you cannot write a sensible program) - maybe we should just require the code capability to be normal.}
-
-\subsubsection{Contexts}
-\begin{definition}
-  \label{def:context}
-  A context, $\context$, is a register file $\reg$, registers $r_1$ and $r_2$, and memory segments $\ms$ and $\ms_\link$ satisfying
-  \begin{itemize}
-  \item $\reg$ respects write-xor-execute with slack $[r_1,r_2]$.
-  \item $\reg(\rstk) = ((\rw,\linear),\stkb,\eaddr,\eaddr)$ where $\stkb \leq \eaddr$\\
-        {\sourcecolor{}$\reg(\rstk) = \stkptr{\rw,\stkb,\eaddr,\eaddr}$ where $\stkb \leq \eaddr$}
-  \item $\ms \uplus \ms_\link$ respects write-xor-execute
-  \item $\ms \uplus \ms_\link$ and $[\reg(\RegName)]$ are self-contained.
-  \item $\ms \uplus \ms_\link$ contains no stack pointer.
-  \item $\reg$ contains no stack pointer with slack $[r_1,r_2]$.
-  \item {\color{red}TODO: No return pointer}
-  \end{itemize}
-In this definition, the source level specific conditions replace the equivalent target level conditions.
-\end{definition}
-
-\subsubsection{Plugging a context}
 
@@
   \]
-  \lau{maybe the last bit should have its own definition.}
   % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
@@
 \begin{definition}[Use return seals only for call]
-  \label{def:use-return-seals-call}
   A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
@@
 \begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
-  \label{def:handle-trusted-seals-properly}
-  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
 \begin{definition}[Never store link seals and code capability]
-  \label{def:never-store-seal-code-cap}
-  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  A configuration $\src{\Phi}$ is said to not store seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
   for all $i \leq n$, if
@@
 \[
-  \Phi'(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
+  \Phi'(r_2) = \seal(\sigma_\baddr,\sigma_\eaddr,\_) \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
 \]
@@
 \begin{definition}
-  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  A program $\program$ with memory segment $\ms_\program$ and $\dom(\ms_\program) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
   \begin{itemize}
-  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-  \item $\plug{\context}{\program}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\plug{\context}{\program}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+  \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
+  \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
+  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
   \end{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex

@@
 \begin{definition}
-  A program $\program$ with memory segment $\ms_\program$ and $\dom(\ms_\program) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
   \begin{itemize}
-  \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
-  \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
-  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
+  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\plug{\context}{\program}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\plug{\context}{\program}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
   \end{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex
reverted notation change and adapted other part.

@@
   \[
-    \npair{\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
-      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray} \in \lreg{\trust}(W)
+    \npair{\left(\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
+      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lreg{\trust}(W)
   \]
@@
 \begin{definition}
-  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
   \begin{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex
First updates to explanatory section about component layout and such.

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
 \subsection{Component layout and linking}
-We define a notion of
-
-A target-level program needs a number of things available in order to be able to execute securely. In the following, we list what it needs followed by a description of how they are made available.
+\dominique{21-3-2018: TODO: update this section (further) to reflect recent changes.}
+A target- or source-level component needs a number of things.
+In the following, we list what it needs followed by a description of how they are made available.
 
@@
 \begin{description}
-\item[Code] The instructions of a program should of course be available so they can be executed.
-\item[Program seals] Each call/return point of the program needs a unique seal in order to enforce well-bracketedness. This seal needs to be unique on the entire machine.
-\item[Stack] A program stores the value of local variables on the stack. 
-\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory separate from the stack.
-\item[Linking table] A program may need to call other programs. Other programs are made available in a linking table.
+\item[Code] The instructions of a component should of course be available so they can be executed.
+\item[Component seals] Each call/return point of the component, as well as every location where new closures are created, needs a unique seal in order to enforce well-bracketedness.
+  This seal needs to be unique on the entire machine.
+\item[Stack] A component stores the value of local variables on the stack. 
+\item[Persistent data storage] A component may also need to use "global variables" these are stored in a piece of memory separate from the stack.
+\item[Access to other components' exported functions] A component may need to call other components.
+  Every component makes available a number of exported values (e.g.
+sealed code-data pairs for invoking certain functions) and components are allowed to include these exported values in their data memory.
+  They could, for example, keep a linking table there through which they can invoke other components' exports.
 \end{description}
 
-In the following, we describe the setup in detail. Figure~\ref{fig:trg-prog-link} illustrates the setup.
-
-The \emph{code}, \emph{program seals}, and \emph{linking table} are available through a normal read-execute capability in the $\pc$ register.
-The program seals are found on the first address governed by this capability.
-The second address governed by the capability contains a normal read capability that governs the linking table and points to the first address.
+In the following, we describe the setup in detail.
+Figure~\ref{fig:trg-prog-link} illustrates the setup (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).
+
+The \emph{code} and \emph{program seals} are available through a normal read-execute capability in the $\pc$ register.
+The program seals are stored as part of the code memory and can be accessed through the $\pc$ register.
+The data memory may contain references to references exported by other components, organised, for example, as a linking table.
 The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
 
-The \emph{linking table} contains pairs of sealed code and data capabilities. So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
-
-The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.} The seals are globally unique in the sense that no other program has access to this range of seals.
+The \emph{linking table} contains pairs of sealed code and data capabilities.
+So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
+
+The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.}
+The seals are globally unique in the sense that no other program has access to this range of seals.
 
@@
   \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
-  \caption{Linking and program layout.}
+  \caption{Linking and program layout (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).}
   \label{fig:trg-prog-link}

### 2018-03-21 fully-abstract-comp/technical-report.tex
Add assumption about seals in components' code pointing to the first seal in the range.

@@
   \inference{
-    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} & [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)
+    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} & [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)
   }{

### 2018-03-21 fully-abstract-comp/technical-report.tex
Change old target macro to be a macro for constructs common between the two languages (that's how it's used) and add a real target macro.
Define initial execution configurations.

@@
 n &\in & \ints \\
-\trg{r} &\in &  \tRegName \\
-\trg{\rn} &\defbnf &  \trg{r} \mid n \\
-\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
-      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
-      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
-  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}} \mid \\ 
-      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
+\com{r} &\in &  \tRegName \\
+\com{\rn} &\defbnf &  \com{r} \mid n \\
+\com{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \\
+      & &  \tgete{\com{r}}{\com{r}}\mid \tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid\\
+      & &  \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
+  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \\ 
+      & &   \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\tsplice{\com{r}}{\com{r}}{\com{r}} 
 \end{array}
@@
     o & \in & \nats \\
-    \src{i} & \defbnf &  \trg{i} \mid \scall{\offpc,\offsigma}{r}{r}
+    \src{i} & \defbnf &  \com{i} \mid \scall{\offpc,\offsigma}{r}{r}
   \end{array}
@@
 \begin{definition}[Programs and Contexts]
-  We define a program as a $(\var{comp},c_{\mathrm{main}})$ component with an empty import list and $c_{\mathrm{main}}$ is one of the exported symbols.
-
-  Given a component $\var{comp}$, a context is a tuple $(\var{comp'},c_{\mathrm{main}})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'},c_{\mathrm{main}})$ is a program.
+  We define a program as a $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ component with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
+
+  Given a component $\var{comp}$, a context is a tuple $(\var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is a program.
+\end{definition}
+
+\begin{definition}[Initial execution configuration]
+  \begin{mathpar}
+    \inference{
+      \reg(\pc) = c_{\mathrm{main},c} & 
+      \reg(\rdata) = c_{\mathrm{main},d} \\
+      \src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
+      \trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
+      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
+      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
+      \overline{c_{\mathrm{import}}} = \emptyset
+    }{
+      ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
+    }
+  \end{mathpar}
 \end{definition}
@@
 \[
-\comp{\cdot} : \src{i}^* \fun \trg{i}^*
+\comp{\cdot} : \src{i}^* \fun \com{i}^*
 \]

### 2018-03-21 fully-abstract-comp/technical-report.tex
Remove some outdated notes.

@@
 \begin{enumerate}
-\item Enter capabilities replaces by sealed code/data pairs \label{item:first-point}
+\item Enter capabilities replaced by sealed code/data pairs \label{item:first-point}
   \begin{itemize}
@@
   \end{itemize}
-\item Components are memory segments and symbols.
-\item Variable instruction length in the source to avoid leaking information through code size that cannot be matched after compilation (breaks compiler correctness, see example in subsection~\ref{subsec:capability-opacity}).
-  \begin{itemize}
-  \item Notice x86 allows instructions of up to infinite length, so this is not weird.
-  \end{itemize}
-\item Have opaque capabilities.
-  \begin{itemize}
-  \item Without opaque capabilities, stack consumption could be inferred through stack pointer index etc.
-  \end{itemize}
-\item Considered PORs for transition systems to: ($a \rightarrow b \Rightarrow a \rightarrow a \text{ and } b \rightarrow b$)
-  \dominique{27-09-2017: I don't remember what we meant with PORs, but perhaps quasi-reflexive partial orders.  Not sure what the point of this was again... }
-  \lau{28-09-2017: I think the next three points is the answer.
-We basically wanted a formalism in which we could design one region that would govern the stack.
-I do not remember why it was made possible with PORs.}
-  \begin{itemize}
-  \item Define one region that governs the stack
-  \item Make the regions more general
-  \item For now, we have not done this: instead we stick with the transition systems we have and have two regions governing the stack reflecting the CSF work.
-  \end{itemize}
 \end{enumerate}

### 2018-03-21 fully-abstract-comp/technical-report.tex
Fix compiler section.

@@
 \label{sec:compiler}
-\[
-\comp{\cdot} : \src{i}^* \fun \com{i}^*
-\]
-\dominique{27-09-2017: Regarding the type signature for $[\cdot]$ above: does this mean that it maps lists of instructions to lists of instructions?
-Is there another function that compiles a single instruction to a list of instructions?}
-\lau{28-09-2017: Yes, that was the thought at the moment, but this may very well change soon.
-This compiler would then use a compiler that compiles single instructions.
-That said, this part was added a long time ago and for some reason (some details about the compilation probably did not make sense to me at the moment), I did not complete it.
-I think the compiler will probably change to be from programs to programs, and a programs is not just going to be a list of instructions.}
-
-\[
-  \comp{\scall{\offpc,\offsigma}{r_1}{r_2}} = 
-  \begin{array}[t]{l}
-    % Make the stack capability into the return capability and seal it:
-    %   Push r_data to the stack (should this be made the responsibility of the programmer?
-If it happens here, then it should probably also be part of the call semantics.
-Especially, the call semantics should "eat" another piece of memory on the stack to make sure there is no discrepancy.)
-    %   Seal the stack capability (with what?
-where does the seal come from?
-I think we talked about this, and it is something that should be handled like linking (a fresh seal is available in the beginning of a programs code along with a capability for a linking table)).
-    % restrict the stack capability to the unused part (and clear this part (is this really necessary?)).
-    % make return capability and ceal it:
-    %   Move a copy of pc to a different register.
-    %   Adjust the pc copy to point to the return address.
-    %   Seal it with the seal.
-    % remove all data from temp registers (that is at least the seal).
-    % jmpx to r1 and r2
-    % restore code:
-    %   check the stack end address.
-    %   move the stack pointer to the proper register
-    %   pop the data capability
-  \end{array}
-\]
-
-\clearpage
+
+The compiler is the identity.
 

### 2018-03-21 fully-abstract-comp/technical-report.tex
Add reference.
Fix compilation.
added precondition to FTLR

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
 n &\in & \ints \\
-\com{r} &\in &  \tRegName \\
-\com{\rn} &\defbnf &  \com{r} \mid n \\
-\com{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \\
-      & &  \tgete{\com{r}}{\com{r}}\mid \tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid\\
-      & &  \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
-  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \\ 
-      & &   \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\tsplice{\com{r}}{\com{r}}{\com{r}} 
+\trg{r} &\in &  \tRegName \\
+\trg{\rn} &\defbnf &  \trg{r} \mid n \\
+\trg{i} &\defbnf & \tfail \mid \thalt \mid \tjmp{\trg{r}} \mid \tjnz{\trg{r}}{\trg{\rn}} \mid \tisptr{\trg{r}}{\trg{r}} \mid \tgeta{\trg{r}}{\trg{r}} \mid \tgetb{\trg{r}}{\trg{r}} \mid \\
+      & &  \tgete{\trg{r}}{\trg{r}}\mid \tgetp{\trg{r}}{\trg{r}} \mid \tgetlin{\trg{r}}{\trg{r}} \mid \tmove{\trg{r}}{\trg{\rn}} \mid \tstore{\trg{r}}{\trg{r}} \mid\\
+      & &  \tload{\trg{r}}{\trg{r}} \mid \tcca{\trg{r}}{\trg{\rn}} \mid \trestrict{\trg{r}}{\trg{\rn}} \mid \tlt{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \\
+  & & \tplus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tminus{\trg{r}}{\trg{\rn}}{\trg{\rn}} \mid \tsetatob{\trg{r}} \mid \txjmp{\trg{r}}{\trg{r}} \mid \tcseal{\trg{r}}{\trg{r}} \mid \\ 
+      & &   \tsplit{\trg{r}}{\trg{r}}{\trg{r}}{\trg{\rn}} \mid\tsplice{\trg{r}}{\trg{r}}{\trg{r}} 
 \end{array}
@@
     o & \in & \nats \\
-    \src{i} & \defbnf &  \com{i} \mid \scall{\offpc,\offsigma}{r}{r}
+    \src{i} & \defbnf &  \trg{i} \mid \scall{\offpc,\offsigma}{r}{r}
   \end{array}
@@
 \subsection{Component layout and linking}
-\dominique{21-3-2018: TODO: update this section (further) to reflect recent changes.}
-A target- or source-level component needs a number of things.
-In the following, we list what it needs followed by a description of how they are made available.
+We define a notion of
+
+A target-level program needs a number of things available in order to be able to execute securely. In the following, we list what it needs followed by a description of how they are made available.
 
@@
 \begin{description}
-\item[Code] The instructions of a component should of course be available so they can be executed.
-\item[Component seals] Each call/return point of the component, as well as every location where new closures are created, needs a unique seal in order to enforce well-bracketedness.
-  This seal needs to be unique on the entire machine.
-\item[Stack] A component stores the value of local variables on the stack. 
-\item[Persistent data storage] A component may also need to use "global variables" these are stored in a piece of memory separate from the stack.
-\item[Access to other components' exported functions] A component may need to call other components.
-  Every component makes available a number of exported values (e.g.
-sealed code-data pairs for invoking certain functions) and components are allowed to include these exported values in their data memory.
-  They could, for example, keep a linking table there through which they can invoke other components' exports.
+\item[Code] The instructions of a program should of course be available so they can be executed.
+\item[Program seals] Each call/return point of the program needs a unique seal in order to enforce well-bracketedness. This seal needs to be unique on the entire machine.
+\item[Stack] A program stores the value of local variables on the stack. 
+\item[Persistent data storage] A program may also need to use "global variables" these are stored in a piece of memory separate from the stack.
+\item[Linking table] A program may need to call other programs. Other programs are made available in a linking table.
 \end{description}
 
-In the following, we describe the setup in detail.
-Figure~\ref{fig:trg-prog-link} illustrates the setup (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).
-
-The \emph{code} and \emph{program seals} are available through a normal read-execute capability in the $\pc$ register.
-The program seals are stored as part of the code memory and can be accessed through the $\pc$ register.
-The data memory may contain references to references exported by other components, organised, for example, as a linking table.
+In the following, we describe the setup in detail. Figure~\ref{fig:trg-prog-link} illustrates the setup.
+
+The \emph{code}, \emph{program seals}, and \emph{linking table} are available through a normal read-execute capability in the $\pc$ register.
+The program seals are found on the first address governed by this capability.
+The second address governed by the capability contains a normal read capability that governs the linking table and points to the first address.
 The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
 
-The \emph{linking table} contains pairs of sealed code and data capabilities.
-So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
-
-The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.}
-The seals are globally unique in the sense that no other program has access to this range of seals.
+The \emph{linking table} contains pairs of sealed code and data capabilities. So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
+
+The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.} The seals are globally unique in the sense that no other program has access to this range of seals.
 
@@
   \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
-  \caption{Linking and program layout (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).}
+  \caption{Linking and program layout.}
   \label{fig:trg-prog-link}
@@
   \inference{
-    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} & [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)
+    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} & [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)
   }{
@@
 \begin{definition}[Programs and Contexts]
-  We define a program as a $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ component with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
-
-  Given a component $\var{comp}$, a context is a tuple $(\var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is a program.
-\end{definition}
-
-\begin{definition}[Initial execution configuration]
-  \begin{mathpar}
-    \inference{
-      \reg(\pc) = c_{\mathrm{main},c} & 
-      \reg(\rdata) = c_{\mathrm{main},d} \\
-      \src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
-      \trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
-      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
-      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
-      \overline{c_{\mathrm{import}}} = \emptyset
-    }{
-      ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
-    }
-  \end{mathpar}
+  We define a program as a $(\var{comp},c_{\mathrm{main}})$ component with an empty import list and $c_{\mathrm{main}}$ is one of the exported symbols.
+
+  Given a component $\var{comp}$, a context is a tuple $(\var{comp'},c_{\mathrm{main}})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'},c_{\mathrm{main}})$ is a program.
 \end{definition}
@@
 \[
-\comp{\cdot} : \src{i}^* \fun \com{i}^*
+\comp{\cdot} : \src{i}^* \fun \trg{i}^*
 \]
@@
     \item $\trust = \trusted$
-    \item the compilation assumptions on $n$, $W$, ...
-    \item the full abstraction conditions (difference with the previous point?)
+
+    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
+      \[
+        \sigrets = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\sigrets',\_} \wedge \sigma \in \sigrets'  \wedge \ta' \subseteq \ta\}
+      \]
+      and
+      \[
+        \sigcloss = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\_,\sigcloss'} \wedge \sigma \in \sigcloss'  \wedge \ta' \subseteq \ta\}
+      \]
+      \begin{itemize}           
+\item $\Phi$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\Phi$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\Phi$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\Phi$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+      \end{itemize}
     \end{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{url}
 
@@
 
+Update: followup paper to the above presented at PriSC 2018, perhaps published elsewhere?
+\url{https://popl18.sigplan.org/event/prisc-2018-formally-secure-compilation-of-unsafe-low-level-components}
+
 \bibliography{references}

### 2018-03-21 fully-abstract-comp/technical-report.tex
removed comment with the old FTLR
added color to FTLR condition

@@
 
-% \begin{theorem}[FTLR]
-%   \label{thm:ftlr}
-%   \[
-%     \begin{array}{l}
-%       \forall n,W,W_D,W_C,\perm_C,\perm_D,\lin_C,\lin_D,\baddr_C,\baddr_D,\eaddr_C,\eaddr_D,\aaddr_C,\aaddr_D \ldotp \\
-%       \quad \left(
-%       \begin{array}{l}
-%         W = W_D \oplus W_C \wedge \\
-%         \perm_D \in \readAllowed{} \Rightarrow \npair{(\baddr_D,\eaddr_D)} \in \readCond{\lin_D,W_D} \wedge\\
-%         \perm_D \in \writeAllowed{} \Rightarrow\npair{(\baddr_D,\eaddr_D)} \in \writeCond{\lin_D,W_D} \wedge\\
-%         \perm_D \notin \{\rwx,\rx\} \wedge \\
-%         \perm_C = \rx \wedge\\
-%         \npair{(\baddr_C,\eaddr_C)} \in \readCond{\lin_C,W_C}
-%       \end{array}
-%       \right)\\
-%       \quad\Rightarrow 
-%       \npair{\left(\arraycolsep=0pt\array{l}((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D),\\
-%                       ((\perm_C,\lin_C),\baddr_C,\eaddr_C,\aaddr_C),((\perm_D,\lin_D),\baddr_D,\eaddr_D,\aaddr_D)\endarray \right)} \in \lre(W)
-
-%     \end{array}
-%   \]
-% \end{theorem}
-
-% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
-It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
-% Note: the below FTLR is simpler than the above thanks to the new expression relation relating pairs of caps rather than pairs of pairs of caps.
-It expresses that any reasonable executable capability is safe to use as the pc capability.
 \begin{theorem}[FTLR]
@@
     \item $\trust = \trusted$
-
     \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
@@
       \begin{itemize}           
-\item $\Phi$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-  \item $\Phi$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\Phi$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-  \item $\Phi$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+\item $\src{\Phi}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\src{\Phi}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\src{\Phi}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\src{\Phi}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
       \end{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex
Forgot two requirements about initial execution configuration.
Small layout change.
have code reg use code region
Small fix.
Fix typo.
Remove outdated comment.
small fix in valid component def
Fix an inconsistency.
Fix problem pointed out by Lau in component validity judgement.
Slight modifications in full abstraction statement.
Remove outdated unfinished definition.

@@
   }{
-    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
   }
@@
     \mscode(a) \in \ints\\
-    \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge \sigma_\baddr+\offsigma \in \sigrets)
+    \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
   }{
-    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
   }
@@
     \forall a \in \dom(\mscode) \ldotp 
-    d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+    \sigrets,d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
   }{
@@
       \trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
+      \reg(\RegName \setminus \{\pc,\rdata,\rstk\}) = 0\\
+      \range{\ms_\stk} = \{0\}&
       \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, \pwheap}, \ms \ldotp\\
+      \exists r \in \addressable{\lin, \pwheap}, \ta, \sigrets, \sigcloss, \code,  \ldotp\\
       \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
-      \quad \pwheap(r) \nequal \stareg[(\ms,\ms)]{\pur}
+      \quad \pwheap(r) \nequal \codereg{\ta,\sigrets,\sigcloss,\code}
     \end{array}
@@
 \[
-  \codereg{\ta,\sigrets,\sigcloss} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigrets \; \sigcloss \; \code)
+  \codereg{\ta,\sigrets,\sigcloss,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigrets \; \sigcloss \; \code)
 \]
@@
       \[
-        \sigrets = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\sigrets',\_} \wedge \sigma \in \sigrets'  \wedge \ta' \subseteq \ta\}
+        \sigrets = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\sigrets',\_,\_} \wedge \sigma \in \sigrets'  \wedge \ta' \subseteq \ta\}
       \]
@@
       \[
-        \sigcloss = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\_,\sigcloss'} \wedge \sigma \in \sigcloss'  \wedge \ta' \subseteq \ta\}
+        \sigcloss = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\_,\sigcloss',\_} \wedge \sigma \in \sigcloss'  \wedge \ta' \subseteq \ta\}
       \]
@@
   \]
-  \dominique{6-3-2018: I added $\xReadCond{\lin,W}$ since we are assuming there are no rwx pointers anywhere.}
-  \begin{definition}[The compilation assumptions]
-    No call uses $\rtmp{1}$ for jumping.\\
-    ...
-  \end{definition}
-
 \end{theorem}
@@
   \]
-  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
+  and $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
 \[
@@
 \[
-  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
+  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
 \]
@@
   \label{thm:full-abstraction}
-  For reasonable programs $\program_1$ and $\program_2$, we have
+  For reasonable components $\program_1$ and $\program_2$, we have
   \begin{gather*}

### 2018-03-21 fully-abstract-comp/technical-report.tex
Remove ta as argument of code region.
Use a macro where useful.

@@
 \begin{align*}
-  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
+  \var{comp} &\mathrel{::=} (\mscode,\msdata,\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
 \end{align*}
@@
 \[
-  \codereg{\ta,\sigrets,\sigcloss,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \ta \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \ta \; \sigrets \; \sigcloss \; \code)
+  \codereg{\sigrets,\sigcloss,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code)
 \]
@@
 \begin{equation*}
-  H^\mathrm{code} \; \ta \; \sigrets \; \sigcloss \; \code \; s \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; s \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
     \dom(\code) = [\aaddr,\baddr]
@@
 \begin{multline*}
-  H^\mathrm{code}_\sigma \; \ta \; \sigrets \; \sigcloss \; \code \; s \; \sigma \; \hat{W} = \\
+  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; s \; \sigma \; \hat{W} = \\
   \begin{array}[t]{l}
@@
 \begin{definition}
-  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  A component $\program = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
   \begin{itemize}

### 2018-03-21 fully-abstract-comp/technical-report.tex
made sure linearity is guaranteed in valid component

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
 \begin{align*}
-  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
+  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
 \end{align*}
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
   }
@@
   \inference{
-    [\baddr,\eaddr] \subseteq A_{\mathrm{data}} & \permbnf \sqsubseteq \rw&
-    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}}
+    \permbnf \sqsubseteq \rw&
+    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}} &
+    \lin = \linear \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
   }
@@
   \inference{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
     \sigma \in \sigcloss
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
+    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
   }
@@
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
-    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),\dom(\msdata),A_{\mathrm{own}}(a),\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash \msdata(a)\\
-    \overline{\dom(\mscode), \dom(\msdata), \emptyset, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
+    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
+    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),\dom(\msdata),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
+    \overline{\dom(\mscode), \dom(\msdata), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
   }{
@@
 \begin{definition}[Programs and Contexts]
-  We define a program as a $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ component with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
+  We define a program $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ as a component with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
 
@@
     \inference{
-      \reg(\pc) = c_{\mathrm{main},c} & 
+      \reg(\pcreg) = c_{\mathrm{main},c} & 
       \reg(\rdata) = c_{\mathrm{main},d} \\
@@
       \trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
-      \reg(\RegName \setminus \{\pc,\rdata,\rstk\}) = 0\\
+      \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0\\
       \range{\ms_\stk} = \{0\}&
@@
   \]
+  \begin{definition}[The compilation assumptions]
+    No call uses $\rtmp{1}$ for jumping.\\
+    ...
+  \end{definition}
+
 \end{theorem}
@@
   \label{thm:full-abstraction}
-  For reasonable components $\program_1$ and $\program_2$, we have
+  For reasonable programs $\program_1$ and $\program_2$, we have
   \begin{gather*}

### 2018-03-21 fully-abstract-comp/technical-report.tex
Some refactoring in FTLR formulation.

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, \pwheap}, \ta, \sigrets, \sigcloss, \code,  \ldotp\\
+      \exists r \in \addressable{\lin, \pwheap}, \sigrets, \sigcloss, \code,  \ldotp\\
       \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
-      \quad \pwheap(r) \nequal \codereg{\ta,\sigrets,\sigcloss,\code}
+      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code}
     \end{array}
@@
 
+\begin{definition}[World specifies component]
+  We define that a world $W$ specifies the presence of a component
+  \begin{equation*}
+    \var{comp} = (\mscode,\msdata,\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
+  \end{equation*}
+  (written as $W \vdash \var{comp}$)
+  iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
+  \begin{itemize}
+  \item $\pwheap(r_i) = \codereg{\sigrets_i,\sigcloss_i,\code_i}$
+  \item $\mscode = \biguplus_i \code_i$
+  \item $\sigrets = \biguplus_i \sigrets_i$
+  \item $\sigcloss = \biguplus_i \sigcloss_i$.
+\end{itemize}
+\end{definition}
+
 \begin{theorem}[FTLR]
@@
   \item \begin{itemize}
-    \item $[\baddr,\eaddr] \subseteq \ta$
+    \item $W \vdash (\mscode,\_,\_,\_,\sigrets,\sigcloss)$
+    \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
     \item $\trust = \trusted$
-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
-      \[
-        \sigrets = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\sigrets',\_,\_} \wedge \sigma \in \sigrets'  \wedge \ta' \subseteq \ta\}
-      \]
-      and
-      \[
-        \sigcloss = \bigcup_{r \in \dom(\pwheap)} \{\sigma \mid \pwheap(r) = \codereg{\ta',\_,\sigcloss',\_} \wedge \sigma \in \sigcloss'  \wedge \ta' \subseteq \ta\}
-      \]
       \begin{itemize}           

### 2018-03-21 fully-abstract-comp/technical-report.tex
removed obsolete value in component judgement

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
   }
@@
   \inference{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
     \sigma \in \sigcloss
   }{
-    A_{\mathrm{code}},A_{\mathrm{data}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
   }
@@
     \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),\dom(\msdata),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
-    \overline{\dom(\mscode), \dom(\msdata), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
+    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
+    \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
   }{

### 2018-03-21 fully-abstract-comp/technical-report.tex
Fix wrongly deleted line.

@@
 \begin{align*}
-  \var{comp} &\mathrel{::=} (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+  \var{comp} &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
 \end{align*}
@@
 \subsection{Fundamental Theorem of Logical Relations}
+
+\begin{definition}[World specifies component]
+  We define that a world $W$ specifies the presence of a component
+  \begin{equation*}
+    \var{comp} = (\mscode,\msdata,\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
+  \end{equation*}
+  (written as $W \vdash \var{comp}$)
+  iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
+  \begin{itemize}
+  \item $\pwheap(r_i) = \codereg{\sigrets_i,\sigcloss_i,\code_i}$
+  \item $\mscode = \biguplus_i \code_i$
+  \item $\sigrets = \biguplus_i \sigrets_i$
+  \item $\sigcloss = \biguplus_i \sigcloss_i$.
+\end{itemize}
+\end{definition}
 
@@
   \]
-  \begin{definition}[The compilation assumptions]
-    No call uses $\rtmp{1}$ for jumping.\\
-    ...
-  \end{definition}
-
 \end{theorem}

### 2018-03-22 fully-abstract-comp/technical-report.tex
Add missing requirement.

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
   (written as $W \vdash \var{comp}$)
-  iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
+  iff
+  \begin{equation*}
+    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+  \end{equation*}
+  and there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
   \begin{itemize}
@@
     \item $\trust = \trusted$
+-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
       \begin{itemize}           

### 2018-03-22 fully-abstract-comp/technical-report.tex
added TODOs

--- 2018-03-21 fully-abstract-comp/technical-report.tex
@@
 \]
+\lau{TODO: Make this condition require a set of adjacent regions. Check whether needed for the stack condition.}
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
@@
 \]
+\lau{TODO: Make this condition require a set of adjacent regions.}
+\lau{TODO: allow code regions to be merged into one. Allow for }
 \dominique{9-3-2018: I changed the above to require that $\ms$ contains safe values by requiring that it is in Hstd. Is this reasonable?}
@@
 \]
+\lau{TODO: Make this condition require a set of adjacent regions. Check whether necessary for the stack region.}
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
@@
   \begin{equation*}
-    \var{comp} = (\mscode,\msdata,\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)
+    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
   \end{equation*}
@@
     \item $\trust = \trusted$
+    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
       \begin{itemize}           

### 2018-03-22 fully-abstract-comp/ftlr-proof-cases.tex
xjmp case was missing in proposed refactoring.
changes

@@
   \end{itemize}
-\item Other cases: store, load, jmp, jnz
+\item Other cases: store, load, jmp, jnz, xjmp
 \end{itemize}
@@
 \end{proof}
+
+\begin{lemma}[call lemma]
+
+\end{lemma}
+
+\section{TODO/Notes}
+\begin{itemize}
+\item Boundary crossing can occur on jump, xjump, call \emph{and} step!
+  \begin{itemize}
+  \item Make a lemma for updatePc that handles the latter once and for all. Maybe make a general lemma for all boundary crossings
+    \begin{itemize}
+    \item updatePc special case
+    \item argue about boundary crossing, trusted/untrusted, sanity conditions
+    \end{itemize}
+  \end{itemize}
+\item Lemma about call
+\end{itemize}
 \end{document}

### 2018-03-22 fully-abstract-comp/technical-report.tex
changes
Remove evolvability of regions (unused).

@@
   \Regionh = 
-  \{\pure \} \times \State \times \Rels \times\\
-  (\State \fun \World \monnefun \URel{\MemSeg^2}) \times \\
-  (\State \fun \Seal \parfun \World \monnefun \URel{\SealableCaps \times \SealableCaps})
+  \{\pure \} \times (\World \monnefun \URel{\MemSeg^2}) \times \\
+  (\Seal \parfun \World \monnefun \URel{\SealableCaps \times \SealableCaps})
 \end{multline*}
@@
   \begin{array}{l}
-    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2}) \uplus \\
-    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2})\uplus \\ 
+    \{\spatial \} \times (\World \monnefun \URel{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times (\World \monnefun \URel{\MemSeg^2})\uplus \\ 
     \{\revoked\}
@@
 
-\[
-  \Rels = \{(\phi_\pub, \phi) \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
-\]
+% \[
+%   \Rels = \{(\phi_\pub, \phi) \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
+% \]
 
@@
 \begin{align*}
-  (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
-  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
-  (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
-                                           =  & \; (\spatialo,\var{sm})
+  (\pure,H,H_\sigma) \oplus (\pure,H,H_\sigma) =  & \; (\pure,H,H_\sigma) \\
+  (\spatial,H) \oplus (\spatial,H) =  & \; (\spatial,H) \\
+  (\spatialo,H) \oplus (\spatial,H) = & \; (\spatial,H) \oplus (\spatialo,H)\\
+                                           =  & \; (\spatialo,H)
 \end{align*}
@@
 Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region.
-It is further not defined if the two sides do not agree on region type or state machine.
-Here $\var{sm}$ is the state machine of the region.
+It is further not defined if the two sides do not agree on region type or heap or sealed value relations.
 
@@
   \begin{array}{l}
-    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) > 0 \wedge \dom(\ms_T) > 0 
+    \forall n, \hat{W}\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \;  \hat{W} \ldotp \dom(\src{\ms_S}) > 0 \wedge \dom(\ms_T) > 0 
   \end{array}
@@
 Also: both future region relations allow $\spatial$ regions to become $\spatialo$, which expresses that our system is affine, rather than linear.
+\lau{22-3-2018: need extra rules for regions with seal interpretations}
 \begin{mathpar}
   \inferrule{ }
-  { \revoked \privft (\spatial,\_,\_,\_,\_)}
+  { \revoked \privft (\spatial,\_)}
 \and
@@
             \and
-  \inferrule{ (s,s') \in \phi}
-            { (\spatial,s',\phi_\pub,\phi,H) \privft (\spatialo,s,\phi_\pub,\phi,H)  }
-            \and
-  \inferrule{ (s,s') \in \phi }
-            { (v,s',\phi_\pub,\phi,H) \privft (v,s,\phi_\pub,\phi,H)  }
+            \inferrule{}{(\spatialo,H) \privft (\spatial,H)}
+% \and
+%   \inferrule{ (s,s') \in \phi}
+%             { (\spatialo,s',\phi_\pub,\phi,H) \privft (\spatial,s,\phi_\pub,\phi,H)  }
+%             \and
+%   \inferrule{ (s,s') \in \phi }
+%             { (v,s',\phi_\pub,\phi,H) \privft (v,s,\phi_\pub,\phi,H)  }
 \end{mathpar}
@@
             \and
-  \inferrule{ (s,s') \in \phi_\pub}
-            { (\spatial,s',\phi_\pub,\phi,H) \pubft (\spatialo,s,\phi_\pub,\phi,H)  }
-            \and
-  \inferrule{ (s,s') \in \phi_\pub}
-            { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
+            \inferrule{}{(\spatialo,H) \privft (\spatial,H)}
+  %           \and
+  % \inferrule{ (s,s') \in \phi_\pub}
+  %           { (\spatialo,s',\phi_\pub,\phi,H) \pubft (\spatial,s,\phi_\pub,\phi,H)  }
+  %           \and
+  % \inferrule{ (s,s') \in \phi_\pub}
+  %           { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
 \end{mathpar}
@@
       \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r)) \\
+      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \\
       \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
@@
       \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r))
+      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
     \end{array}
@@
       \qquad \forall r \in \activeReg{\pwheap}, n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \pwheap(r).s \; \xi^{-1}(R_W(r)) \wedge\\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
       \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \text{ is defined }\\
-      \quad \dom(\pwheap(r).H_\sigma \; \pwheap(r).s) = R_\var{seal}(r)
+      \quad \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
     \end{array}
@@
 %       \qquad \forall r \in \dom(W) , n' < n \ldotp \\
-%       \quad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
+%       \quad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (\xi^{-1} (R_w (r))) 
 %     \end{array}
@@
       \qquad \exists r \in \addressable{l,S(i)} \ldotp\\
-      \qquad \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; S(i)(r).s \; \xi^{-1}(S(i))  \ldotp \\
+      \qquad \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; \xi^{-1}(S(i))  \ldotp \\
       \qquad \qquad \dom(\src{\ms_S}) \supseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \supseteq [\baddr,\eaddr]      
@@
       \npair{(\src{\vsc_S},\vsc_T)} \in \lrp(W) \wedge \\
-      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand \\
-      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; s \; \sigma \; W\wedge \\
-      \quad (\isLinear{\vsc_1} \Rightarrow \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
+      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
+      \quad (\isLinear{\vsc_1} \Rightarrow \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
       \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-      \quad (\nonLinear{\vsc_1} \Rightarrow \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; s \; \sigma \; W_o \ldotp \\
+      \quad (\nonLinear{\vsc_1} \Rightarrow \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
       \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
@@
     \; \middle| \;
-    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,s,\_,\_,\_,H_\sigma) \tand H_\sigma\; s \; \sigma \nequal \lrv
+    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma \nequal \lrv
     \right\} \cup \\
@@
   \label{def:address-stratified}
-  We say that a region $\iota = (\_,\_,\_,\_,H)$ is address stratified iff
+  We say that a region $\iota = (\_,H,\_)$ is address stratified iff
   \[
@@
       \forall n, \src{\ms_S},\ms_T,\src{\ms_S'},\ms_T',s,\hat{W}\ldotp \\
-      \quad \npair{\stpair{\ms}{\ms}}, \npair{\stpair[.]{\ms_S'}{\ms_T'}} \in H \; s \; \hat{W} \wedge \\
+      \quad \npair{\stpair{\ms}{\ms}}, \npair{\stpair[.]{\ms_S'}{\ms_T'}} \in H \; \hat{W} \wedge \\
       \quad \dom(\src{\ms_S}) = \dom(\ms_T) = \dom(\src{\ms_S'}) = \dom(\ms_T') \\
       \quad \Rightarrow \\
-      \qquad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\src{\ms_S}\update{\aaddr}{\src{\ms_S'}(\aaddr)},\ms_T\update{\aaddr}{\ms_T'(\aaddr)})} \in H \; s \; \hat{W}
+      \qquad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\src{\ms_S}\update{\aaddr}{\src{\ms_S'}(\aaddr)},\ms_T\update{\aaddr}{\ms_T'(\aaddr)})} \in H \; \hat{W}
     \end{array}
@@
 \[
-  \stdreg{A}{v} \defeq (v,*,=,=,H_A^{\mathrm{std}})
+  \stdreg{A}{v} \defeq (v,H_A^{\mathrm{std}})
 \]
@@
 \[
-  H_A^{\mathrm{std}} \; * \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
+  H_A^{\mathrm{std}} \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
     \begin{array}{l}
@@
 \[
-  \stareg{v} = (v,*,=,=,H^\mathrm{sta}_{\stpair{\ms}{\ms}})
+  \stareg{v} = (v,H^\mathrm{sta}_{\stpair{\ms}{\ms}})
 \]
@@
 \[
-  \codereg{\sigrets,\sigcloss,\code} \defeq (\pure,*,=,=,H^\mathrm{code} \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code)
+  \codereg{\sigrets,\sigcloss,\code} \defeq (\pure,H^\mathrm{code} \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code)
 \]
@@
 \begin{equation*}
-  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; s \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
     \dom(\code) = [\aaddr,\baddr]
@@
 \begin{multline*}
-  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; s \; \sigma \; \hat{W} = \\
+  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \sigma \; \hat{W} = \\
   \begin{array}[t]{l}
@@
       \begin{array}{l}
-        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, ((\rx,\lin),\baddr',\eaddr',\aaddr'+\calllen)) \tand \\
+        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, ((\rx,\linear),\baddr',\eaddr',\aaddr'+\calllen)) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
@@
 \end{theorem}
+\begin{proof}
+  First, 
+  TODO
+\end{proof}
 

### 2018-03-22 fully-abstract-comp/technical-report.tex
memsat non-empty stack frames and higher stack frames have lower addresses.
naming
memSat fix, frame 0 highest frame
Remove the public/private distinction in the future world relation, as per a discussion here in Leuven with Lau.

@@
 % {\color{DarkGreen}
-The public part of the future world relation stays the same:
+The future world relation becomes:
 \begin{mathpar}
-  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \pubft W.i(r) }
+  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r) }
             { W' \future W }
 \end{mathpar}
-The private future world relation becomes:
-\begin{mathpar}
-  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \privft W.i(r) \\  \exists r \in \dom(W'.\mathrm{priv}) \setminus \dom(m_{\mathrm{priv}}(\pwpriv))\ldotp \pwpriv[W'](r) \text{ is non-empty}}
-            { W' \future W }
-\end{mathpar}
-where we take ``is non-empty'' to mean the following:
-
-\begin{definition}
-  We say a region $(k,s,\phi_\pub,\phi,H)$ is non-empty iff
-\[
-  \begin{array}{l}
-    \forall n, \hat{W}\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \;  \hat{W} \ldotp \dom(\src{\ms_S}) > 0 \wedge \dom(\ms_T) > 0 
-  \end{array}
-\]
-\end{definition}
+% The private future world relation becomes:
+% \begin{mathpar}
+%   \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \privft W.i(r) \\  \exists r \in \dom(W'.\mathrm{priv}) \setminus \dom(m_{\mathrm{priv}}(\pwpriv))\ldotp \pwpriv[W'](r) \text{ is non-empty}}
+%             { W' \future W }
+% \end{mathpar}
+% where we take ``is non-empty'' to mean the following:
+
+% \begin{definition}
+%   We say a region $(k,s,\phi_\pub,\phi,H)$ is non-empty iff
+% \[
+%   \begin{array}{l}
+%     \forall n, \hat{W}\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \;  \hat{W} \ldotp \dom(\src{\ms_S}) > 0 \wedge \dom(\ms_T) > 0 
+%   \end{array}
+% \]
+% \end{definition}
 
@@
 %
-The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$.
-Also: both future region relations allow $\spatial$ regions to become $\spatialo$, which expresses that our system is affine, rather than linear.
+Future regions allow $\spatial$ regions to become $\revoked$.
+Also: the future region relation allows $\spatial$ regions to become $\spatialo$, which expresses that our system is affine, rather than linear.
 \lau{22-3-2018: need extra rules for regions with seal interpretations}
 \begin{mathpar}
-  \inferrule{ }
-  { \revoked \privft (\spatial,\_)}
+  \inferrule{ }{ \revoked \future (\spatial,\_)}
 \and
-  \inferrule{ }
-            { \revoked \privft \revoked }
+  \inferrule{ }{ \revoked \future \revoked }
   \and
-            \inferrule{}{(\spatialo,H) \privft (\spatial,H)}
+  \inferrule{}{(\spatialo,H) \future (\spatial,H)}
+  \and
+  \inferrule{}{(\spatialo,H) \future (\spatialo,H)}
+  \and
+  \inferrule{}{(\spatial,H) \future (\spatial,H)}
 % \and
@@
 %
-Public future regions are allowed to transition according to the public part of their transition system.
-\begin{mathpar}
-    \inferrule{ }
-            { \revoked \pubft \revoked }
-            \and
-            \inferrule{}{(\spatialo,H) \privft (\spatial,H)}
+% Public future regions are allowed to transition according to the public part of their transition system.
+% \begin{mathpar}
+%     \inferrule{ }
+%             { \revoked \pubft \revoked }
 %             \and
-  % \inferrule{ (s,s') \in \phi_\pub}
-  %           { (\spatialo,s',\phi_\pub,\phi,H) \pubft (\spatial,s,\phi_\pub,\phi,H)  }
-  %           \and
-  % \inferrule{ (s,s') \in \phi_\pub}
-  %           { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
-\end{mathpar}
+%             \inferrule{}{(\spatialo,H) \privft (\spatial,H)}
+%   %           \and
+%   % \inferrule{ (s,s') \in \phi_\pub}
+%   %           { (\spatialo,s',\phi_\pub,\phi,H) \pubft (\spatial,s,\phi_\pub,\phi,H)  }
+%   %           \and
+%   % \inferrule{ (s,s') \in \phi_\pub}
+%   %           { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
+% \end{mathpar}
 
@@
       \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
+      \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
+      \quad \forall j < i \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp a < a') \wedge\\
       \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\

### 2018-03-22 fully-abstract-comp/technical-report.tex
code region requires code to be valid wrt seals. ret seals empty in code region when code not subset of Ta
step bug
typos
changed opc on call stack to just be return address rather than capability
Removed permission and linearity from return code pointer
cleaned up the call semantics
updated priv stack region to require address instead of cap + typos
naming

@@
     & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
-    & & & {\sourcecolor{} \; \mid \retptrd(\basebnf,\aendbnf) \mid \retptrc(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf)}\\
+    & & & {\sourcecolor{} \; \mid \retptrd(\basebnf,\aendbnf) \mid \retptrc(\basebnf,\aendbnf,\addrbnf)}\\
     \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}\\ 
@@
     &\MemSeg & \defeq & \Addr \parfun \Word \\
-    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg}\\
+    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Addr \times \MemSeg}\\
     \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
@@
     \src{\baddr} & \src{\text{if $\vsc = \retptrd(\baddr,\_)$}} \\ 
-    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\_,\_,\baddr,\_,\_)$}}
+    \src{\baddr} & \src{\text{if $\vsc = \retptrc(\baddr,\_,\_)$}}
   \end{cases}
@@
     \src{\eaddr} & \src{\text{if $\vsc = \retptrd(\_,\eaddr)$}} \\ 
-    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,\_,\_,\eaddr,\_)$}}
+    \src{\eaddr} & \src{\text{if $\vsc = \retptrc(\_,\eaddr,\_)$}}
   \end{cases}
@@
     \src{\linear} & \src{\text{if $\vsc = \retptrd(\_,\_)$}} \\ 
-    \src{\lin} & \src{\text{if $\vsc = \retptrc(\_,\lin,\_,\_,\_)$}}
+    \src{\normal} & \src{\text{if $\vsc = \retptrc(\_,\_,\_)$}}
   \end{cases}
@@
       c = \retptrd(\_,\_)\\
-      c = \retptrc(\_,\linear,\_,\_,\_)
     \end{array}\right.\\
@@
                                                     \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
-                                                    \lbrack \aaddr ,\baddr \rbrack \subseteq \ta \tand \\
+                                                    \lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta \tand \\
                                                     \baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr \tand \\
@@
 \[
-\encType{}_{\var{src}}(\src{\retptrd(\_,\_)}) = \encType{}_{\var{src}}(\src{\retptrc(\_,\_,\_,\_,\_)}) = \encType{}_{\var{trg}}(((\_,\_),\_,\_,\_))
+\encType{}_{\var{src}}(\src{\retptrd(\_,\_)}) = \encType{}_{\var{src}}(\src{\retptrc(\_,\_,\_)}) = \encType{}_{\var{trg}}(((\_,\_),\_,\_,\_))
 \]
@@
         \Phi' & \updReg{r_2}{0} \\
-              & \updReg{\pcreg}{\opc}\\
+              & \updReg{\pcreg}{c_\opc}\\
               & \updReg{\rdata}{0}\\
@@
       \begin{array}[t]{l}
-        c_1 = \retptrc(\_) \tand \\
+        c_1 = \retptrc(\baddr,\eaddr,\aaddr) \tand \\
         c_2 = \retptrd(\aaddr_\stk,\eaddr_{\stk,\priv}) \tand \\
@@
         \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
+        \opc = \aaddr \tand \\
+        c_\opc = ((\rx,\normal),\baddr,\eaddr,\opc) \tand \\
         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
@@
                                                         &\updReg{\rstk}{c_\stk}\\
-                                                        &\left[
-\arraycolsep=0pt\array{l}
-\reg.\rretc \mapsto \\
-\hfill \sealed{\sigma_?,\retptrc(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen)}
-\endarray
-\right]\\
-                                                        &\updReg{\rretd}{\sealed{\sigma_?,\retptrd(\aaddr_\stk,\eaddr_\stk)}}\\
+                                                        &\updReg{\rretc}{\sealed{\sigma_?,c_\opc}}\\
+                                                        &\updReg{\rretd}{\sealed{\sigma_?,c_\var{priv\_data}}}\\
                                                         &\updReg{\rtmp{1}}{0}
@@
                                                    c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
-                                                   \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                                   \opc = ((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen) \tand \\
+                                                   c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) \tand \\
+                                                   \Phi(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \tand \\
+                                                   \opc = \aaddr + \calllen \tand \\
+                                                   c_\opc = \retptrc(\baddr,\eaddr,\aaddr+\calllen) \tand \\
                                                    \stk' = (\opc,\ms_{\stk,\priv}) :: \stk \tand\\
@@
     \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
-  }
+  }b
   \and
@@
 \[
-  \Worlds = \RegionName \parfun (\Regions \times \SealableCaps)
+  \Worlds = \RegionName \parfun (\Regions \times \Addr)
 \]
@@
   \and
-  \inferrule{}{(\spatialo,H) \future (\spatial,H)}
+  \inferrule{ }{(\spatialo,H) \future (\spatial,H)}
   \and
-  \inferrule{}{(\spatialo,H) \future (\spatialo,H)}
+  \inferrule{ }{(\spatialo,H) \future (\spatialo,H)}
   \and
-  \inferrule{}{(\spatial,H) \future (\spatial,H)}
+  \inferrule{ }{(\spatial,H) \future (\spatial,H)}
 % \and
@@
                        \begin{cases}
-                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\sigma)\\
-                         ((\spatial,\var{sm}),\sigma) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\sigma)\\
-                         (\revoked,\sigma) & \text{if } W_\var{priv}(r) = (\revoked,\sigma)\\
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\opc)\\
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\opc)\\
+                         (\revoked,\opc) & \text{if } W_\var{priv}(r) = (\revoked,\opc)\\
                        \end{cases}\\
@@
       \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
-      \quad \forall j < i \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp a < a') \wedge\\
+      \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp a < a') \wedge\\
       \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
@@
   H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
-    \dom(\code) = [\aaddr,\baddr]
+    \begin{array}{l}
+    \dom(\code) = [\aaddr,\baddr] \wedge \\
+    \dom(\code) \nsubseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
+    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
+    \end{array}
   \right\}
@@
 \lau{19-03-2018: Add the link capability to the above? code in ta?}
+\lau{TODO: look at the $[b,e] \subseteq ...$ condition - will it work for splice?}
 \begin{multline*}
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\rx,\lin,\baddr,\eaddr,\aaddr'+\calllen - 1),\\((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
       \\
@@
         \aaddr = \aaddr' + \retoffset \tand \\
-        \code(a+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
+        \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
         \lbrack \baddr, \eaddr \rbrack \subseteq \dom(\code) \subseteq \ta
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
       \\
       \begin{array}{l}
-        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, ((\rx,\linear),\baddr',\eaddr',\aaddr'+\calllen)) \tand \\
+        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
         \quad \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \quad \code(a+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
-        \quad a = b-1\tand \\
+        \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
         \quad \lbrack \baddr', \eaddr' \rbrack \subseteq \dom(\code) \subseteq \ta

### 2018-03-23 fully-abstract-comp/technical-report.tex
removed P (for now)
Remove two TODOs in the value relation.

@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}},\stpair[.]{v_{d,T}}{v_{d,S}}]} \in \lrp(W) \wedge \\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{[\stpair[.]{v_{c,S}}{v_{c,T}}]} \in \lrp(W) \wedge \\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
@@
     \begin{array}{l}
-      \npair{(\src{\vsc_S},\vsc_T)} \in \lrp(W) \wedge \\
       \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
@@
     \begin{array}{l}
-      \npair{(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
       \begin{array}{r l l}
+        \perm = \noperm & \Rightarrow & \npair{[(\baddr,\eaddr)]} \in \lrp(W) \wedge \\
         \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
@@
     \begin{array}{l}
-      \npair{(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrp(W) \wedge \\
       \begin{array}{r l l }
+        \perm = \noperm & \Rightarrow & \npair{[(\baddr,\eaddr)]} \in \lrp(W) \wedge \\
         \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
@@
     \; \middle| \;
-    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \text{TODO: code region for trusted code}
+    \begin{array}{l}
+      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \codereg{\sigrets,\sigcloss,\code} \wedge\\
+      \hfill\dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)
+    \end{array}
     \right\} \cup \\
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
-    \begin{array}{l}
-      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \text{TODO: code region for trusted code} \wedge\\
-      \perm \sqsubseteq \rx \wedge [\baddr,\eaddr] \subseteq \dom(\var{code})
-    \end{array}\right\}
-  \end{array}
-\]
+      \perm \sqsubseteq \rx \wedge \npair{(\baddr,\eaddr)} \in \xReadCond{\normal,W}
+    \right\}
+  \end{array}
+\]
+
+Note: the case for sub-$\rx$ capabilities in the trusted value relation allows for pointers to trusted code blocks.
+Such pointers will not satisfy the read condition, which requires the standard region, which is defined in terms of the untrusted value relation.
+Trusted code blocks contain trusted seal capabilities which do not satisfy the untrusted code relation.
+An alternative might be to introduce a trust parameter to the readcondition and standard region and merge the case with the regular case for rx capabilities.
 
@@
       \quad \forall n' < n, W' \future W_l, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
-      \quad \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}]} \in \lrp(W') \Rightarrow \\
-      \qquad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}} \in \lre(W')
+      \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}} \in \lre(W')
     \end{array}

### 2018-03-23 fully-abstract-comp/technical-report.tex
reformulate a condition.
updated H sigma requirements on range
fix compilation.
removed range requirement in H sigma in the data ret case for the code pointer
typos
Small fixes here and there.
Fix a problem in the FTLR statement related to the readcondition.
Define contextual equivalence.

@@
                                                     \lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta \tand \\
-                                                    \baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr \tand \\
+                                                    \lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr] \tand \\
                                                     \exec{\Phi(\pcreg)}
@@
     \tgetb{\rtmp{1}}{\rstk}\\
-    \tminus{\rtmp{1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
+    \tminus{\rtmp{1}}{\rtmp{1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
     \tmove{\rtmp{2}}{\pcreg}\\
@@
       \begin{array}[t]{rl}
-        \Phi' & \updReg{r_2}{0} \\
-              & \updReg{\pcreg}{c_\opc}\\
+        \Phi' & \updReg{\pcreg}{c_\opc}\\
               & \updReg{\rdata}{0}\\
@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\}) (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
 % Lau 19-01-2018: Added rdata as default.
-
+\lau{TODO check what conditions need to have a ``square'' i.e. need approximation}
 \[
@@
       \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
-      \quad (\isLinear{\vsc_1} \Rightarrow \forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
-      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-      \quad (\nonLinear{\vsc_1} \Rightarrow \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; W_o \ldotp \\
-      \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
+      \quad (\isLinear{\vsc_S} \Rightarrow \\
+      \qquad\forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+      \quad (\nonLinear{\vsc_S} \Rightarrow \\
+      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
 
@@
       \begin{array}{l}
-        \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
         \aaddr = \aaddr' + \retoffset \tand \\
         \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
-        \lbrack \baddr, \eaddr \rbrack \subseteq \dom(\code) \subseteq \ta
+        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
       \end{array}
@@
       \begin{array}{l}
-        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, \aaddr'+\calllen) \tand \\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
-        \quad \decInstr{\code([a',a' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
         \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
-        \quad \lbrack \baddr', \eaddr' \rbrack \subseteq \dom(\code) \subseteq \ta
       \end{array}
@@
         \sigma \in \sigcloss \tand\\
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; W) \tor\\
-        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; W)
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
+        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
       \end{array}
@@
   \begin{itemize}
-  \item $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
+  \item $\trust = \trusted \vee \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
   \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
@@
 \end{theorem}
+Note: we don't require the readcondition in the trusted case because trusted code pointers point to trusted code blocks which may require seals for trusted seals which are not in the untrusted value relation.
 
@@
 \begin{definition}
-  A component $\program = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  A component $\program = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
   \begin{itemize}
-  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-  \item $\plug{\context}{\program}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\plug{\context}{\program}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $A$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\plug{\context}{\program}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\plug{\context}{\program}$ handle trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
   \end{itemize}
+\end{definition}
+
+\begin{definition}[Source language contextual equivalence]
+  In the source language, we define that $\src{\program_1 \sconeq \program_2}$ iff
+  \begin{equation*}
+    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \term^{\ta_1}} \Leftrightarrow \src{\plug{\context}{\program_2} \term^{\ta_2}}
+  \end{equation*}
+  with $\src{\ta_i} = \src{\dom(\program_i.\mscode)}$
+\end{definition}
+
+\begin{definition}[Target language contextual equivalence]
+  In the target language, we define that $\trg{\program_1 \tconeq \program_2}$ iff
+  \begin{equation*}
+    \forall \trg{\context} \ldotp \trg{\plug{\context}{\program_1} \term} \Leftrightarrow \trg{\plug{\context}{\program_2} \term}
+  \end{equation*}
 \end{definition}

### 2018-03-23 fully-abstract-comp/technical-report.tex
updated static region to require values in untrusted value relation
fixed compilation error
Explanation about source language contextual equivalence.

@@
 
+Note that we define source language contextual equivalence with respect to contexts that are not in $\ta$.
+This means that they are unable to perform calls.
+We believe this fits with the goal of this work: allow programmers (or better: authors of previous compiler passes) to reason about their target language programs under a special perspective, where all calls can be interpreted as calls that actually behave in a well-bracketed way by the operational semantics.
+This special perspective is defined by overlaying a different operational semantics on the existing code: the source language semantics.
+The fact that source contexts cannot make calls themselves is no problem: authors of previous compiler passes should only be able to take the perspective that their own calls are guaranteed to be well-bracketed.
+It does not matter for them whether calls in the rest of the system are guaranteed to be well-bracketed.
+Note also that if the trusted code hands out closures, the context can still invoke them with an xjmp, rather than a call.
+That xjmp can even be the one in the implementation of call if the context uses that implementation. 
+This works perfectly fine, except that the context does not get any well-bracketedness guarantees, but that doesn't matter.
+
 \begin{definition}[Target language contextual equivalence]

### 2018-03-23 fully-abstract-comp/technical-report.tex
Made T_A parameter of termination condition. Added TODO to add this to the rest of source semantics.
Added TODOs from last day in Leuven

--- 2018-03-23 fully-abstract-comp/technical-report.tex
@@
 We use the following step relation:
-
+\lau{TODO: Add $\ta$ to the source step.}
 \begin{align*}
@@
   \Phi\term \defeq \exists i \ldotp \term[i]
+\]
+We use a similar notation for source configuration termination
+\[
+  \Phi\sterm[i]{\ta} \defeq \Phi \nstep[i] \halted
+\]
+if we just know $\Phi$ terminates in some number of steps, then we write
+\[
+  \Phi\sterm{\ta} \defeq \exists i \ldotp \sterm[i]{\ta}
 \]
@@
 % {\color{DarkGreen}
+\lau{TODO add frame to memory satisfaction and remove it from observation relation.}
 Memory satisfaction for new worlds:
@@
       \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
-      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term[i] \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
+      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta} \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
     \end{array}
@@
       \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
-      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term
+      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term{\ta}
     \end{array}
@@
 \[
-  H^\mathrm{sta}_{\stpair{\ms}{\ms}} = \{ \npair{\stpair{\ms}{\ms}} \}
+  H^\mathrm{sta}_{\stpair{\ms}{\ms}} = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+    \begin{array}{l}
+      \exists S : \dom(\ms) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
+      \quad \forall \aaddr \in \dom(\ms) \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+    \end{array}
+\right\}
 \]
@@
 \section{Full Abstraction}
+\lau{TODOs from last day in Leuven}
+\lau{TODO: components have optional main (pc/data pair)}
+\lau{TODO: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
+\lau{TODO: Component logical relation.}
+\lau{TODO: Version of FTLR for programs.}
+\lau{TODO: Component lemma for initial executable configurations.}
+\lau{TODO: Executable configuration lemma.}
+\lau{TODO: ... should use initial executable configuration.}
 \begin{definition}
@@
   \begin{equation*}
-    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \term^{\ta_1}} \Leftrightarrow \src{\plug{\context}{\program_2} \term^{\ta_2}}
+    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \sterm{\ta[,1]}} \Leftrightarrow \src{\plug{\context}{\program_2} \sterm{\ta[,2]}}
   \end{equation*}
-  with $\src{\ta_i} = \src{\dom(\program_i.\mscode)}$
+  with $\src{\ta[,i]} = \src{\dom(\program_i.\mscode)}$
 \end{definition}

### 2018-03-23 fully-abstract-comp/technical-report.tex
added TODO for base_stack parameter
began passing TA and stk_base around in gc record (may have missed some places).

@@
 \lau{TODO: Add $\ta$ to the source step.}
-\begin{align*}
-  \src{\Phi} & \; \src{\step \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
+\lau{TODO: Add $\stkb$ to source term and everything else where it is necessary.
+Consider how $\stkb$ and $\ta$ can be added without passing around these values in LR and source semantics (solution to parameterize both with it, so they need to be provided only once.)}
+\begin{align*}
+  \src{\Phi} & \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} &  &\sourcecolor\left.\arraycolsep=0pt
                                                   \begin{array}[t]{l}
@@
                                                   \end{array}\right.\\
-  \Phi & \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
+  \Phi & \step[\src{\ta,\stkb}] \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) & &\left.\arraycolsep=0pt
                                                   \begin{array}[t]{l}
@@
                                                   \end{array}\right.\\
-  \Phi& \step \failed & & \totherwise
+  \Phi& \step[\src{\ta,\stkb}] \failed & & \totherwise
 \end{align*}
@@
 \[
-  \Phi\term[i] \defeq \Phi \nstep[i] \halted
+  \Phi\sterm[i]{\src{\ta,\stkb}} \defeq \Phi \nstep[i]{\src{\ta,\stkb}} \halted
 \]
@@
 \[
-  \Phi\term \defeq \exists i \ldotp \term[i]
-\]
-We use a similar notation for source configuration termination
-\[
-  \Phi\sterm[i]{\ta} \defeq \Phi \nstep[i] \halted
-\]
-if we just know $\Phi$ terminates in some number of steps, then we write
-\[
-  \Phi\sterm{\ta} \defeq \exists i \ldotp \sterm[i]{\ta}
+  \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}
 \]
@@
       \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
-      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta} \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
+      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
     \end{array}
@@
       \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
-      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \term{\ta}
+      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
     \end{array}
@@
     \begin{array}{l}
-      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \codereg{\sigrets,\sigcloss,\code} \wedge\\
+      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \codereg{\sigrets,\sigcloss,\code,\gc} \wedge\\
       \hfill\dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)
@@
       \quad \exists A \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwheap(r) \nsubeq \stdreg{A}{\pur}
+      \qquad \pwheap(r) \nsubeq \stdreg{A,\gc}{\pur}
     \end{array}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwfree(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}
+      \qquad \pwfree(r) \nsubeq \stdreg{[\baddr',\eaddr'],\gc}{\pur}
     \end{array}
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, \pwheap}, \sigrets, \sigcloss, \code,  \ldotp\\
+      \exists r \in \addressable{\lin, \pwheap}, \sigrets, \sigcloss, \code  \ldotp\\
       \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
-      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code}
+      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}
     \end{array}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
+      \qquad \pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr'],\gc}{\pur} \wedge \\
       \qquad \pwheap(r) \text{ is address-stratified}
@@
 \dominique{Why $n-1$ above?
-There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr']}{\pur}$.
+There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr'],\gc}{\pur}$.
 Also: readCond doesn't have the $n-1$.}
@@
       \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwfree(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur} \wedge \\
+      \qquad \pwfree(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr'],\gc}{\pur} \wedge \\
       \qquad \pwfree(r) \text{ is address-stratified}
@@
 \[
-  \stdreg{A}{v} \defeq (v,H_A^{\mathrm{std}})
+  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std}} \; \gc)
 \]
@@
 \[
-  H_A^{\mathrm{std}} \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
+  H_A^{\mathrm{std}} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
     \begin{array}{l}
@@
 \[
-  \stareg{v} = (v,H^\mathrm{sta}_{\stpair{\ms}{\ms}})
-\]
-\[
-  H^\mathrm{sta}_{\stpair{\ms}{\ms}} = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+  \stareg[\stpair{\ms}{\ms},\gc]{v} = (v,H^\mathrm{sta}_{\stpair{\ms}{\ms}} \; \gc)
+\]
+\[
+  H^\mathrm{sta}_{\stpair{\ms}{\ms}} \; \gc = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
     \begin{array}{l}
@@
 \[
-  \codereg{\sigrets,\sigcloss,\code} \defeq (\pure,H^\mathrm{code} \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code)
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,H^\mathrm{code} \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
 \]
@@
 \begin{multline*}
-  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \sigma \; \hat{W} = \\
+  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \hat{W} = \\
   \begin{array}[t]{l}
@@
       \begin{array}{l}
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms)]{\spa}, \aaddr'+\calllen) \tand \\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms),\gc]{\spa}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
@@
 \begin{definition}[World specifies component]
-  We define that a world $W$ specifies the presence of a component
+  We define that a world $W$ specifies the presence of a component with global constants $\gc$
   \begin{equation*}
@@
   \end{equation*}
-  (written as $W \vdash \var{comp}$)
+  (written as $W \vdash_\gc \var{comp}$)
   iff
@@
   \begin{itemize}
-  \item $\pwheap(r_i) = \codereg{\sigrets_i,\sigcloss_i,\code_i}$
+  \item $\pwheap(r_i) = \codereg{\sigrets_i,\sigcloss_i,\code_i,\gc}$
   \item $\mscode = \biguplus_i \code_i$
@@
   \item \begin{itemize}
-    \item $W \vdash (\mscode,\_,\_,\_,\sigrets,\sigcloss)$
+    \item $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$
     \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
@@
   \[
-    \src{\Phi} \nstep[i] \src{\Phi'}
+    \src{\Phi} \nstep[i]{(\ta,\stkb)} \src{\Phi'}
   \]
@@
   % \[
-  %   \src{\Phi} \nstep[j] \src{\Phi''}
+  %   \src{\Phi} \nstep[j]{} \src{\Phi''}
   % \]
@@
   \[
-    \src{\Phi} \nstep[i] \src{\Phi'}
+    \src{\Phi} \nstep[i]{(\ta,\_)} \src{\Phi'}
   \]
@@
   \[
-    \src{\Phi} \nstep[i] \src{\Phi'} \step \src{\Phi''}
+    \src{\Phi} \nstep[i]{(\ta,\stkb)} \src{\Phi'} \step[(\ta,\stkb)] \src{\Phi''}
   \]
@@
   \begin{equation*}
-    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \sterm{\ta[,1]}} \Leftrightarrow \src{\plug{\context}{\program_2} \sterm{\ta[,2]}}
+    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\program_2} \sterm{\ta[,2],\stkb_2}}
   \end{equation*}

### 2018-03-23 fully-abstract-comp/technical-report.tex
Added frame to memSat instead of lro.
added boxed to permission based conditions and updated a few mistakes. Added TODO

@@
 We use the following step relation:
-\lau{TODO: Add $\ta$ to the source step.}
-\lau{TODO: Add $\stkb$ to source term and everything else where it is necessary.
+\lau{RESOLVED: Add $\ta$ to the source step.}
+\lau{RESOLVED: Add $\stkb$ to source term and everything else where it is necessary.
 Consider how $\stkb$ and $\ta$ can be added without passing around these values in LR and source semantics (solution to parameterize both with it, so they need to be provided only once.)}
@@
 % {\color{DarkGreen}
-\lau{TODO add frame to memory satisfaction and remove it from observation relation.}
+\lau{RESOLVED add frame to memory satisfaction and remove it from observation relation.}
 Memory satisfaction for new worlds:
@@
     \begin{array}{l}
+      \exists  \ms_T' \ldotp \\
+
       \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
@@
       W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
-      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
-      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
+      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\ldotp \\
+      \quad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
+      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
+
       \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
       \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
-      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W_{\var{heap}}}
+      \quad \memSatHeap{\ms_S',\ms_\var{T,heap}}{W_{\var{heap}}}
     \end{array}
@@
     \begin{array}{l}
-      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
-      \quad \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \Rightarrow (\ms_T \uplus \ms_{f,T},\reg_T) \term\\
+      \forall i \leq n \ldotp \\
+      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \Rightarrow (\ms_T,\reg_T) \term\\
     \end{array}
@@
     \begin{array}{l}
-      \forall i \leq n, \src{\ms_{f,S}}, \ms_{f,T} \ldotp \\
-      \quad (\ms_T \uplus \ms_{f,T},\reg_T) \term[i] \Rightarrow \src{(\ms_S \uplus \ms_{f,S},\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
+      \forall i \leq n \ldotp \\ 
+      \quad (\ms_T ,\reg_T) \term[i] \Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
     \end{array}
@@
 
+\lau{TODO make this only take a pair (the other uses are gone)}
 \[
@@
 \[
-  \lrvtrusted(W) =
+  \lrvtrusted[\square,(\ta,\stkb)](W) =
   \begin{array}[t]{l}
@@
     \begin{array}{l}
-      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \codereg{\sigrets,\sigcloss,\code,\gc} \wedge\\
+      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
       \hfill\dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)
@@
     \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
-      \perm \sqsubseteq \rx \wedge \npair{(\baddr,\eaddr)} \in \xReadCond{\normal,W}
+    \begin{array}{l}
+      \perm \sqsubseteq \rx \wedge \\
+      \npair{(\baddr,\eaddr)} \in \xReadCond[\square,(\ta,\stkb)]{\normal,W}
+    \end{array}
     \right\}
@@
 \[
-  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std}} \; \gc)
+  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc)
 \]
@@
 \[
-  H_A^{\mathrm{std}} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
+  H_A^{\mathrm{std},\square} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
     \begin{array}{l}
@@
 \[
-  \stareg[\stpair{\ms}{\ms},\gc]{v} = (v,H^\mathrm{sta}_{\stpair{\ms}{\ms}} \; \gc)
-\]
-\[
-  H^\mathrm{sta}_{\stpair{\ms}{\ms}} \; \gc = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+  \stareg[\stpair{\ms}{\ms},\gc]{v,\square} = (v,H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc)
+\]
+\[
+  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
     \begin{array}{l}
@@
 \[
-  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,H^\mathrm{code} \; \sigrets \; \sigcloss \; \code,H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
+H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc,
+H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
 \]
@@
 \begin{equation*}
-  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
     \begin{array}{l}
@@
 \begin{multline*}
-  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \hat{W} = \\
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \; \hat{W} = \\
   \begin{array}[t]{l}
@@
       \begin{array}{l}
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms),\gc]{\spa}, \aaddr'+\calllen) \tand \\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
@@
 \lau{TODO: ... should use initial executable configuration.}
+\lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}
 \begin{definition}

### 2018-03-26 fully-abstract-comp/technical-report.tex
typos
updated P relation to only take a few parameters

@@
       \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \\
+      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
@@
       \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
-      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \text{ is defined }\\
+      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \text{ is defined } \wedge\\
       \quad \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
@@
 
-\lau{TODO make this only take a pair (the other uses are gone)}
-\[
-  \lrp(W) =
-    \left\{
-    \npair{ [\stpair[.]{c_1}{c'_1}, \dots, \stpair[.]{c_k}{c'_k}] } \middle| 
-    \begin{array}{l}
-      \exists S : \{1,\dots,k\} \fun \World \ldotp \\
-      \quad W = \bigoplus_{i\in\{1,\dots,k\}} S(i)\wedge \\
-      \quad\text{for $i \in \{1,\dots,k\}$} \\
-      \qquad\stpair[.]{c_i}{c'_i} \in \ints \times \ints \vee \\
-      \qquad\src{\baddr} = \pi_\baddr(\src{c_i}) \wedge \src{\eaddr} = \pi_\eaddr(\src{c_i}) \wedge \\
-      \qquad\baddr = \pi_\baddr(c_i') \wedge \eaddr = \pi_\eaddr(c_i') \wedge \\
-      \qquad\src{\lin} = \pi_\lin(\src{c_i}) = \pi_\lin(c_i') \wedge \\
-      \qquad \exists r \in \addressable{l,S(i)} \ldotp\\
-      \qquad \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; \xi^{-1}(S(i))  \ldotp \\
-      \qquad \qquad \dom(\src{\ms_S}) \supseteq [\src{\baddr},\src{\eaddr}] \wedge \dom(\ms_T) \supseteq [\baddr,\eaddr]      
+\lau{RESOLVED make this only take a pair (the other uses are gone)}
+\[
+  \lrp(W) = \left\{ \npair{(\lin,\baddr,\eaddr)} \middle|
+    \begin{array}{l}
+      \exists r \in \addressable{l,W} \ldotp\\
+      \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; \xi^{-1}(W)  \ldotp \\
+      \qquad \dom(\src{\ms_S}) \supseteq [\baddr,\eaddr] \wedge \dom(\ms_T) \supseteq [\baddr,\eaddr]
     \end{array}
@@
 % Lau 19-01-2018: Added rdata as default.
-\lau{TODO check what conditions need to have a ``square'' i.e. need approximation}
+\lau{RESOLVED check what conditions need to have a ``square'' i.e. need approximation}
 \[
@@
       \begin{array}{r l l}
-        \perm = \noperm & \Rightarrow & \npair{[(\baddr,\eaddr)]} \in \lrp(W) \wedge \\
+        \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
         \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
@@
       \begin{array}{r l l }
-        \perm = \noperm & \Rightarrow & \npair{[(\baddr,\eaddr)]} \in \lrp(W) \wedge \\
+        \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
         \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
@@
 \lau{TODO: Make this condition require a set of adjacent regions.}
-\lau{TODO: allow code regions to be merged into one. Allow for }
+\lau{TODO: allow code regions to be merged into one. Allow for (... forgot to write the rest, but probably related to splicing regions.) }
 \dominique{9-3-2018: I changed the above to require that $\ms$ contains safe values by requiring that it is in Hstd. Is this reasonable?}

### 2018-04-03 fully-abstract-comp/technical-report.tex
Modify xjump expression relation as discussed with Lau in Leuven.
removed step taken in memory satisfaction. we did not have it in local cap. work, so we probably don't need it here.
typos

@@
     \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}} &
-    \lin = \linear \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
+    \lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
   }{
@@
 Also: the future region relation allows $\spatial$ regions to become $\spatialo$, which expresses that our system is affine, rather than linear.
-\lau{22-3-2018: need extra rules for regions with seal interpretations}
+\lau{22-3-2018: TODO need extra rules for regions with seal interpretations}
 \begin{mathpar}
@@
 
+\begin{lemma}[purePart is idempotent]
+  \label{lem:purePart-idempotemt}
+  For all $W$, we have that $\purePart{W} = \purePart{W} \oplus \purePart{W}$ and $\purePart{W} = \purePart{}(\purePart{W})$.
+\end{lemma}
+
 \begin{lemma}[purePart is duplicable]
@@
     \begin{array}{l}
-      \exists  \ms_T' \ldotp \\
-
       \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
@@
       \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
-      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}})\ldotp \\
+      \qquad \quad \npair{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
@@
       \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
-      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}) \ldotp \\
+      \qquad \quad \npair{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
     \end{array}
@@
       \qquad W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
-      \qquad \forall r \in \activeReg{\pwheap}, n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \qquad \forall r \in \activeReg{\pwheap} \ldotp \\
+      \qquad \quad \npair{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
@@
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\}) (W_\lrrs ) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
-      \qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S}\\
-      \qquad\Phi_T = (\ms_T,\reg_T)\\
-      \qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\\
-      \qquad W \oplus W_\lrrs \oplus W_\lrm \\
-      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
+      \qquad\Phi_T = (\ms_T,\reg_T) \wedge\\
+      \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp\\
+      \quad\qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\\
+      \quad\qquad\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
     \end{array}
@@
       \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
-      \quad (\isLinear{\vsc_S} \Rightarrow \\
+      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
       \qquad\forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-      \quad (\nonLinear{\vsc_S} \Rightarrow \\
+      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
       \qquad \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\

### 2018-04-04 fully-abstract-comp/technical-report.tex
Proposed change to seal case in untrusted value relation.
Requirement on all seals in range of available seals rather than on the specific seal pointed to.
added a suggestion for permission-based conditions using adjacent regions.
proposed update to execCond to handle subseg operation
resolved TODO about stack base address in memsat and main optional

@@
 \begin{definition}[Programs and Contexts]
-  We define a program $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ as a component with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
-
-  Given a component $\var{comp}$, a context is a tuple $(\var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is a program.
+  We define a program to be either a component $\var{comp}$ or a tuple $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with component $\comp$ with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
+
+A context is a program defined as above.
+% Given a component $\var{comp}$, a context is a tuple $(\var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is a program.
+
+Two programs $\var{comp}$ and $(\var{comp}', c_{\mathrm{main},c}, c_{\mathrm{main},d})$ can be linked into a new program $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
 \end{definition}
@@
       \quad \ms_\stk = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \stkb \in \dom(\ms_T) \wedge \stkb \in \dom(\ms_\stk) \wedge \\
       \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
@@
     \; \middle| \;
-    \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma \nequal \lrv
+    \begin{array}{l}
+      \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp\pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal \lrv
+    \end{array}
     \right\} \cup \\
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, \pwheap} \ldotp \\
-      \quad \exists A \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwheap(r) \nsubeq \stdreg{A,\gc}{\pur}
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
+      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
 \]
-\lau{TODO: Make this condition require a set of adjacent regions. Check whether needed for the stack condition.}
+\lau{RESOLVED: Make this condition require a set of adjacent regions. Check whether needed for the stack condition.}
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
@@
     \begin{array}{l}
-      \exists r \in \addressable{\linear, \pwfree} \ldotp \\
-      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwfree(r) \nsubeq \stdreg{[\baddr',\eaddr'],\gc}{\pur}
+      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
+      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, \pwheap}, \sigrets, \sigcloss, \code  \ldotp\\
-      \quad \dom(\ms) \supseteq [\baddr,\eaddr] \wedge \\
-      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}
+      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
+      \quad \exists M : S \fun \powerset{\MemSeg}, R : S \fun \powerset{\Seal}, C : S \fun \powerset{\Seal} \\
+      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
+      \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
+      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{R(r),C(r),M(r),\gc}
     \end{array}
@@
 \]
-\lau{TODO: Make this condition require a set of adjacent regions.}
+\lau{RESOLVED: Make this condition require a set of adjacent regions.}
 \lau{TODO: allow code regions to be merged into one. Allow for (... forgot to write the rest, but probably related to splicing regions.) }
@@
     \begin{array}{l}
-      \exists r \in \addressable{\lin, \pwheap} \ldotp \\
-      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr'],\gc}{\pur} \wedge \\
-      \qquad \pwheap(r) \text{ is address-stratified}
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
+      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq[n-1] \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \quad \pwheap(r) \text{ is address-stratified}
     \end{array}
@@
     \begin{array}{l}
-      \exists r \in \addressable{\linear, \pwfree} \ldotp \\
-      \quad \exists [\baddr',\eaddr'] \supseteq [\baddr,\eaddr] \ldotp \\
-      \qquad \pwfree(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr'],\gc}{\pur} \wedge \\
-      \qquad \pwfree(r) \text{ is address-stratified}
+      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
+      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq[n-1] \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \quad \pwfree(r) \text{ is address-stratified}
     \end{array}
@@
       \quad ((W_l = \purePart{W} \wedge \lin = \normal) \vee (W_l = W \wedge \lin = \linear)) \wedge\\
-      \quad \forall n' < n, W' \future W_l, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
-      \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)}} \in \lre(W')
+      \quad \forall n' < n, W' \future W_l, \perm \in S_\perm, \aaddr \in [\baddr',\eaddr'] \subseteq [\baddr,\eaddr] \ldotp\\
+      \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr',\eaddr',\aaddr)}{((\perm,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
     \end{array}
@@
 \]
-\lau{01-11-2017: TODO look into what kind of world separation should happen here.}
-\dominique{7-3-2018: is it necessary to have $n-1$ above? there is already a quantification over $n'<n$ in the memory satisfaction relation.}
+\lau{01-11-2017: RESOLVED look into what kind of world separation should happen here.}
+%\dominique{7-3-2018: is it necessary to have $n-1$ above?
+there is already a quantification over $n'<n$ in the memory satisfaction relation.} \lau{04-04-2018: The n'<n was removed from memory satisfaction.}
 
@@
 \lau{TODOs from last day in Leuven}
-\lau{TODO: components have optional main (pc/data pair)}
-\lau{TODO: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
+\lau{RESOLVED: components have optional main (pc/data pair)}
+\lau{RESOLVED?: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
 \lau{TODO: Component logical relation.}

### 2018-04-06 fully-abstract-comp/technical-report.tex
Change quantification order in the value relation's new seal case so that it can support splicing of seals.
Remove outdated comment.
Remove n-1 in writecondition to mirror readcondition and hopefully fix a problem I had in a proof.
Typo.

@@
     \begin{array}{l}
-      \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp\pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal \lrv
+      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal \lrv
     \end{array}
@@
 \lau{TODO: allow code regions to be merged into one. Allow for (... forgot to write the rest, but probably related to splicing regions.) }
-\dominique{9-3-2018: I changed the above to require that $\ms$ contains safe values by requiring that it is in Hstd. Is this reasonable?}
 
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
-      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq[n-1] \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
-\dominique{Why $n-1$ above?
+% \dominique{Why $n-1$ above?
 There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr'],\gc}{\pur}$.
-Also: readCond doesn't have the $n-1$.}
+Also: readCond doesn't have the $n-1$.
+Update: the memory satisfaction was updated by Lau to remove the quantification over $n' < n$ and I dropped the $n-1$ here.}
 \[
@@
 \lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
-where $H_A$ is defined as follows:
+where $H^{\mathrm{std},\square}_A$ is defined as follows:
 \[

### 2018-04-06 fully-abstract-comp/technical-report.tex
Fix a TODO, probably a transcription error caused by my illegible handwriting :).
Fix more transcription errors in the TODOs.

@@
 \lau{TODO: Version of FTLR for programs.}
-\lau{TODO: Component lemma for initial executable configurations.}
-\lau{TODO: Executable configuration lemma.}
-\lau{TODO: ... should use initial executable configuration.}
+\lau{TODO: Compatibility lemma for initial executable configuration construct, i.e. a lemma that says ``if two programs are related, then their initial executable configurations will be related''.}
+\lau{TODO: A logical relation for executable configurations.}
+\lau{TODO: contextual equivalence should be defined in terms of initial executable configuration.}
 \lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}

### 2018-04-06 fully-abstract-comp/technical-report.tex
Add a note about definitions of programs, components, contexts.

@@
 \end{definition}
+\dominique{A program is intuitively a component that is ready to be run.
+From that point of view, I don't see the point of a program without main.
+I would propose changing the definitions: a program should be a component that is ready to be executed, i.e.
+it must have an empty import list and a pair of capabilities to be used as main.
+A component can be either a component with a main (i.e.
+a program that still needs to be linked with library implementations) or one without (i.e.
+a library implementation that will be used by other components).
+A context for a given component is any other component that can be linked with it to produce a program.}
 

### 2018-04-07 fully-abstract-comp/technical-report.tex
More notes from black board notes.
Implement proposed redefinition of components, programs, contexts.

--- 2018-04-06 fully-abstract-comp/technical-report.tex
@@
 
+A component can be either a component with a main (i.e.
+a program that still needs to be linked with library implementations) or one without (i.e.
+a library implementation that will be used by other components).
+It contains code memory, data memory a list of imported symbols, a list of exported symbols, a list of seals used for producing return capability pairs and a list of seals used for producing closures.
+
 We define a component as follows:
 \begin{align*}
-  \var{comp} &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+  \var{comp}_0 &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)\\
+  \var{comp} &\mathrel{::=} \var{comp}_0\\
+               &\mid  (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
 \end{align*}
@@
     \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
-  }b
+  }
   \and
@@
   }
+  \and
+  \inference{
+    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)\\
+    \vdash \var{comp}_0 & c_{\mathrm{main},c}, c_{\mathrm{main},d} \in \overline{c_{\mathrm{export}}}
+  }{
+    \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
+  }
 \end{mathpar}
@@
   }
+  \and
+  \inference{
+    \var{comp}_0'' = \var{comp}_0 \bowtie \var{comp}_0'
+  }{
+    (\var{comp}_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \var{comp}_0 \bowtie (\var{comp}_0',c_{\mathrm{main},c}, c_{\mathrm{main},d})
+  }
+  \and
+  \inference{
+    \var{comp}_0'' = \var{comp}_0 \bowtie \var{comp}_0'
+  }{
+    (\var{comp}_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \var{comp}_0'
+  }
 \end{mathpar}
@@
 
+A program is intuitively a component that is ready to be executed, i.e. it must have an empty import list and a pair of capabilities to be used as main.
+A context for a given component is any other component that can be linked with it to produce a program.
+
 \begin{definition}[Programs and Contexts]
-  We define a program to be either a component $\var{comp}$ or a tuple $(\var{comp}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with component $\comp$ with an empty import list and $c_{\mathrm{main},c}$ and $c_{\mathrm{main},d}$ are two of the exported symbols.
-
-A context is a program defined as above.
-% Given a component $\var{comp}$, a context is a tuple $(\var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is a program.
-
-Two programs $\var{comp}$ and $(\var{comp}', c_{\mathrm{main},c}, c_{\mathrm{main},d})$ can be linked into a new program $(\var{comp} \bowtie \var{comp'}, c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
+  We define a program to be a component $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
+
+  A context for a component $\var{comp}$ is another component $\var{comp}'$ such that $\var{comp} \bowtie \var{comp}'$ is a program.
 \end{definition}
@@
 \lau{RESOLVED?: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
-\lau{TODO: Component logical relation.}
+\lau{TODO: Component logical relation. Rough sketch is in white-board-notes/leuven-23-3-2018-3.jpg (top left).}
+\dominique{TODO: Compatibility lemma for linking: if $c_1$ and $c_1'$ are related and $c_2$ and $c_2'$, then so should be $c_1\bowtie c_2$ and $c_1' \bowtie c_2'$ (if defined).}
 \lau{TODO: Version of FTLR for programs.}

### 2018-04-07 fully-abstract-comp/technical-report.tex
Propose LR for components, compatibility lemma for linking.

@@
   \begin{equation*}
-    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
   \end{equation*}
-  (written as $W \vdash_\gc \var{comp}$)
+  (written as $W \vdash_\gc \var{comp}_0$)
   iff
@@
 
+\subsection{Related components}
+\label{sec:related-components}
+
+\begin{align*}
+  \lrcomp(W) =
+  \left\{\begin{multlined}
+     \npair{\var{comp},\var{comp}} \;\mid \;\\
+      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+      W \vdash \var{comp} \tand\\
+      \forall W' \ldotp \text{If } W \oplus W' \text{ is defined }\\
+      \tand \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(W')\\
+      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(W)
+  \end{multlined}
+    \right\} \cup
+  \left\{\begin{multlined}
+     \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
+     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
+      W \vdash \var{comp} \tand\\
+      \forall W' \ldotp \text{If } W \oplus W' \text{ is defined }\\
+      \tand \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(W')\\
+      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(W)
+  \end{multlined}
+    \right\} \cup
+\end{align*}
+
+\begin{lemma}
+  If
+  \begin{itemize}
+  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$
+  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$
+  \end{itemize}
+  then
+  \begin{itemize}
+  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W_1\oplus W_2)$.
+  \end{itemize}
+\end{lemma}
+
 \section{Full Abstraction}
@@
 \lau{RESOLVED: components have optional main (pc/data pair)}
-\lau{RESOLVED?: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
-\lau{TODO: Component logical relation. Rough sketch is in white-board-notes/leuven-23-3-2018-3.jpg (top left).}
-\dominique{TODO: Compatibility lemma for linking: if $c_1$ and $c_1'$ are related and $c_2$ and $c_2'$, then so should be $c_1\bowtie c_2$ and $c_1' \bowtie c_2'$ (if defined).}
+\lau{RESOLVED: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
+\lau{RESOLVED?: Component logical relation. Rough sketch is in white-board-notes/leuven-23-3-2018-3.jpg (top left).}
+\dominique{RESOLVED?: Compatibility lemma for linking: if $c_1$ and $c_1'$ are related and $c_2$ and $c_2'$, then so should be $c_1\bowtie c_2$ and $c_1' \bowtie c_2'$ (if defined).}
 \lau{TODO: Version of FTLR for programs.}

### 2018-04-10 presentations/esop18-04-2018/main.tex
ESOP presentation (work in progress)

--- 2017-12-19 presentations/illustrations/attacks.tex
@@
+\documentclass{beamer}
+\addtobeamertemplate{navigation symbols}{}{%
+    \usebeamerfont{footline}%
+    \usebeamercolor[fg]{footline}%
+    \hspace{1em}%
+    \insertframenumber/\inserttotalframenumber
+}
+
+\input{../illustrations/preamble}
+\title{Reasoning About a Machine with Local Capabilities}
+\subtitle{Provably Safe Stack and Return Pointer Management}
+\author{\textbf{Lau Skorstengaard}\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
+\institute{\inst{1} Aarhus University \and %
+  \inst{2} imec-DistriNet, KU Leuven}
+
+\date{ESOP, April 17, 2018}
+
+\tikzset{
+    gray on/.style={alt=#1{gray}{}},
+    alt/.code args={<#1>#2#3}{%
+      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
+    },
+}
+
+\begin{document}
+\maketitle
+
+\begin{frame}
+  \frametitle{}
+%  let x = ref 0 in
+%    λf.(x := 0; f() x := 1; f(); !x)
+  % Motivation
+\end{frame}
+
+\begin{frame}
+  \frametitle{Traditional Stack Pointers}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
+    \end{scope}
+    \draw (-1.5,0) node {};
+    \draw (-1.5,15) node {};
+    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
+
+    % traditional:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
+
+    % traditional call:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
+
+    % traditional return:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
+
+    % traditional call, attack:
+    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
+    \draw<.-> (8,7) node[draw=black] (sp0) {};
+    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
+
+    % read other stack frames
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
+    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
+
+    % break well-bracketedness
+    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
+    \draw<.-> (8,7) node[draw=black] (sp0) {};
+    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
+    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
+      skip caller frame};
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}
+  % Desired properties
+  % Capability machine as enforcement mechanism
+\end{frame}
+
+\begin{frame}
+  \frametitle{Road map?}
+  % Road map?
+\end{frame}
+
+\begin{frame}
+  \frametitle{Capability machine}
+  % Short description of capabilities
+\end{frame}
+
 \begin{frame}
@@
     \end{scope}
-    \draw (15.5,3) node {heap memory};
+    \draw (14.5,3) node {heap memory};
 
@@
 \end{frame}
-%%% Local Variables:
-%%% TeX-master: "presentation-illustrations"
-%%% End:
+
+\begin{frame}
+  \frametitle{Local capabilities}
+  % Short description of local capabilities
+  \begin{itemize}
+  \item Capabilities tagged with locality (local or normal)
+  \item New \emph{write-local} permission.
+  \item Local capabilities can only be stored by capabilities with \emph{write-local} permission
+  \end{itemize}
+  Calling convention highlights
+  \begin{itemize}
+  \item Stack capability is local with permission read, write-local, and execute.
+  \item Clear stack before passing stack capability to untrusted code.
+  \end{itemize}
+
+\end{frame}
+
+% \begin{frame}
+%   \frametitle{Calling convention}
+%   % and stack management  
+%   \begin{itemize}
+%   \item Initially:
+%     \begin{itemize}
+%     \item Stack capabilitiy local write-local capability
+%     \item No global write-local capabilities on the machine
+%     \end{itemize}
+%   \item Prior to returning to untrusted code:
+%     \begin{itemize}
+%     \item Clear the stack
+%     \item Clear non-return registers
+%     \end{itemize}
+%   \item Prior to calls to untrusted code:
+%     \begin{itemize}
+%     \item Push activation record to stack and create enter-capability for it
+%     \item Restrict the stack pointer to the unused part and clear that part
+%     \item Clear non-argument registers
+%     \item Only invoke global call-backs
+%     \end{itemize}
+%   \item When invoked by untrusted code, make sure stack capability has permission read, write-local and execute.
+%   \end{itemize}
+% \end{frame}
+
+\begin{frame}
+  \frametitle{Local Stack Capabilities Prevent Attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker...
+    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
+    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw node[right=0cm of sp1] { stack pointer };
+
+    % stores stack pointer in heap
+    \begin{scope}
+      \fill[gray!20] (14,6) rectangle (18,13);
+      \draw (14,6) -- (14,13);
+      \draw (18,6) -- (18,13);
+    \end{scope}
+    \draw (16,14) node {heap memory};
+    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
+    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack Clearing Prevents Attack 2}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker stores copy of stack pointer high in the stack
+    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \begin{onlyenv}<+>
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
+    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);
+
+    % attacker calls trusted code again
+    \begin{onlyenv}<+->
+      \inactadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % trusted code clears the stack
+    \begin{onlyenv}<+->
+      \foreach \x in {9,9.5,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
+    (13,13.5) -- (10,14.5);
+
+    % attacker uses old stack pointer
+    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \begin{onlyenv}<.>
+      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % attacker returns
+    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
+    \begin{onlyenv}<.>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % trusted code clears stack again
+    \begin{onlyenv}<+->
+      \foreach \x in {7,7.5,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+    % then, the trusted code can return
+    \begin{onlyenv}<+>
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
+
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Local Stack Capabilities Prevent Attack 3}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    \begin{scope}
+      \fill[gray!50] (13,4) rectangle (18,14);
+      \draw (13,4) -- (13,14);
+      \draw (18,4) -- (18,14);
+      \draw (14.5,15) node {code memory};
+    \end{scope}
+
+    \begin{scope}
+      \fill[gray!50] (13,0) rectangle (18,2);
+      \draw (13,0) -- (13,2);
+      \draw (18,0) -- (18,2);
+      \draw (14.5,3) node {heap memory};
+    \end{scope}
+
+    \begin{onlyenv}<+->
+      \actsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \draw[fill=white] (13,5) rectangle (18,6.5) node[pos=.5] {\footnotesize caller code};
+    \end{onlyenv}
+
+\note{Caller calls callee}
+\note{Caller clears the stack}
+    \begin{onlyenv}<+->
+      \foreach \x in {5.5,6,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+
+\note{Return pointers always point to the stack as we need to be able to regain the stack pointer itself}
+\note{note: the open ended braces are assumed to stretch over the rest of the stack (in the direction it is open)}
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,2.5) rectangle (8.1,5.5);
+        \capbrace[ret1]{(6.1,2.5)}{(6.1,6.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr} };
+      \draw node[right=0cm of ret1] { ret-ptr };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \draw[fill=gray!30] (13,5) rectangle (18,6.5) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,7) rectangle (18,8.5) node[pos=.5] {\footnotesize callee code};
+      \actadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \end{onlyenv}
+
+\note{Callee cannot store ret-ptr on the heap as it is local!}
+    \begin{onlyenv}<+-.(2)>
+      \draw (13,1.5) edge[red,bend right,thick,->] (ret1);
+    \end{onlyenv}
+
+    \begin{onlyenv}<+>
+      \draw[teal,very thick] (11.3,3) -- (13,2.8);
+      \draw[teal,very thick] (11.8,3.7) -- (12.4,2);
+      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,10) {Return pointer is local!};  
+    \end{onlyenv}
+
+\note{Callee tries to store ret-ptr on the stack (just like for attack 2)}
+    \begin{onlyenv}<+->
+      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \end{onlyenv}
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+\note{Callee calls callee (1)}
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,5.5) rectangle (8.1,8.5);
+        \capbrace[ret2]{(6.1,5.5)}{(6.1,9.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr (1)} };
+      \draw node[right=0cm of ret2] { ret-ptr (1) };
+    \end{onlyenv}
+    \draw<.-.(1)> (6,13.25) edge[red,thick,->,in=45] (ret1);
+
+    \begin{onlyenv}<+->
+      % \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      \actsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+      \draw[fill=gray!30] (13,7) rectangle (18,8.5) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,9) rectangle (18,10.5) node[pos=.5] {\footnotesize callee code (1)};
+    \end{onlyenv}
+
+\note{Callee (1) clears the stack erasing the copy of the ret-ptr!}
+    \begin{onlyenv}<+->
+      \foreach \x in {11.5,12,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+
+\note{Callee (1) calls callee (2)}
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \clip (6.1,8.5) rectangle (8.1,11.5);
+        \capbrace[ret3]{(6.1,8.5)}{(6.1,12.5)}
+      \end{scope}
+      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+      \draw node[right=0cm of ret3] { ret-ptr (2) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      % \inactsf{(0,2.5)}{(6,5.5)}{caller stack frame}
+      % \inactadv{(0,5.5)}{(6,8.5)}{callee stack frame}
+      \inactsf{(0,8.5)}{(6,11.5)}{callee stack frame (1)}
+      \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+      \draw[fill=gray!30] (13,9) rectangle (18,10.5) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,11) rectangle (18,12.5) node[pos=.5] {\footnotesize callee code (2)};
+    \end{onlyenv}
+
+\note{The copy of ret-ptr callee (2) hoped to use has been erased}
+    \begin{onlyenv}<+->
+      \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) {ret-ptr capability gone,\\can only access ret-ptr (2)!};  
+    \end{onlyenv}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Contributions}
+  \begin{itemize}
+    \item Formalisation of a capability machine with local capabilities
+    \item Novel calling convention for enforcing control-flow correctness and local-state encapsulation
+    \item A calling convention utilizing local capabilities to ensure well-bracketed control-flow and local-state encapsulation
+    \item Step-indexed Kripke Logical Relation to reason about programs on said machine
+      \begin{itemize}
+        \item Single orthogonal closure
+        \item Variant of public/private future world relation to express nature of local capabilities
+      \end{itemize}
+      %new insights for low-level systems?
+    \item FTLR
+      \begin{itemize}
+      \item Semantic statement of guarantees offered by capability machine
+      \end{itemize}
+    \item Logical relation applied to a number of examples
+      \begin{itemize}
+      \item ``Awkward example''
+      \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\end{document}

### 2018-04-10 fully-abstract-comp/technical-report.tex
Fix layout.
Modify proposed component relation to require purity of imports and exports.
Another update to the component relation
* Quantify over future worlds in which imports may be valid.
* Fix relation requirements for main capability pair
Define context plugging as component linking + initial execution configuration construction.

@@
 
+\begin{definition}[Plugging a program into a context]
+  When $\var{comp'}$ is a context for component $\var{comp}$, $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$.
+  Then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
+\end{definition}
+
 \section{Compiler}
@@
 \begin{align*}
-  \lrcomp(W) =
+  \lrcomp(W) &=
   \left\{\begin{multlined}
@@
       W \vdash \var{comp} \tand\\
-      \forall W' \ldotp \text{If } W \oplus W' \text{ is defined }\\
-      \tand \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(W')\\
-      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(W)
+      \text{For all } W' \future W \ldotp \text{If } \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'})\\
+      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
   \end{multlined}
-    \right\} \cup
-  \left\{\begin{multlined}
+    \right\}\\
+  &\cup \left\{
+    \begin{multlined}
      \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
-     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
-      W \vdash \var{comp} \tand\\
-      \forall W' \ldotp \text{If } W \oplus W' \text{ is defined }\\
-      \tand \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(W')\\
-      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(W)
+     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
+      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+     \{(c_{\mathrm{main},c}, c_{\mathrm{main},c}),(c_{\mathrm{main},d}, c_{\mathrm{main},d})\} \subseteq (\overline{c_{\mathrm{export}}},\overline{c_{\mathrm{export}}})
        \end{multlined}
-    \right\} \cup
+    \right\} 
 \end{align*}

### 2018-04-10 fully-abstract-comp/technical-report.tex
Clarify phrase structure.

@@
 \begin{definition}[Plugging a program into a context]
-  When $\var{comp'}$ is a context for component $\var{comp}$, $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$.
-  Then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
+  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
+  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
 \end{definition}

### 2018-04-10 fully-abstract-comp/technical-report.tex
Some more TODOs.

@@
 
-\begin{lemma}
+\begin{lemma}[Compatibility lemma for linking]
   If
@@
   \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W_1\oplus W_2)$.
+  \end{itemize}
+\end{lemma}
+
+\subsection{Related execution configurations}
+\label{sec:related-exec-confs}
+
+\[
+  \lrec(W) = \left\{ \npair{\left(
+        \arraycolsep=0pt\array{l}
+        (\ms_S,\reg_S,\stk,\ms_\stk),\\
+        (\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
+      \reg_S(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr) \tand (\trust = \trusted \text{ iff } \aaddr \in \ta) \tand \\
+      \npair{( \reg_S(\pcreg),\reg_T(\pcreg) )} \in \lrvg{\trust}(W_\pcreg) \tand \\
+      \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
+      \npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)
+    \end{array}
+            \right\}
+\]
+
+\begin{lemma}[Compatibility lemma for initial execution configuration construction]
+  If
+  \begin{itemize}
+  \item $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$
+  \item $\var{comp}_S\rightsquigarrow \Phi_S$
+  \item $\var{comp}_T \rightsquigarrow \Phi_T$,
+  \end{itemize}
+  then
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W)$
   \end{itemize}
@@
 \lau{TODO: Version of FTLR for programs.}
-\lau{TODO: Compatibility lemma for initial executable configuration construct, i.e. a lemma that says ``if two programs are related, then their initial executable configurations will be related''.}
-\lau{TODO: A logical relation for executable configurations.}
-\lau{TODO: contextual equivalence should be defined in terms of initial executable configuration.}
+\lau{RESOLVED: Compatibility lemma for initial executable configuration construct, i.e.
+a lemma that says ``if two programs are related, then their initial executable configurations will be related''.}
+\lau{RESOLVED: A logical relation for executable configurations.}
+\lau{RESOLVED: contextual equivalence should be defined in terms of initial executable configuration.}
 \lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}

### 2018-04-11 fully-abstract-comp/technical-report.tex
An adequacy result, add parameters to component and exec-conf relations.
Make the "world specifies component" relation deal with untrusted components appropriately.

@@
   \item $\mscode = \biguplus_i \code_i$
-  \item $\sigrets = \biguplus_i \sigrets_i$
-  \item $\sigcloss = \biguplus_i \sigcloss_i$.
-  \end{itemize}
+  \item One of the following holds:
+    \begin{itemize}
+    \item $\dom(\mscode) \subseteq \ta$ and $\sigrets = \biguplus_i \sigrets_i$ and $\sigcloss = \biguplus_i \sigcloss_i$.
+    \item $\dom(\mscode) \mathrel{\#} \ta$ and $\sigrets = \emptyset$ and $\sigcloss = \biguplus_i \sigrets_i \uplus \biguplus_i \sigcloss_i$.
+    \end{itemize}
+  \end{itemize}
+  with $\gc = (\ta,\stkb)$
 \end{definition}
+Note that for untrusted components, we require all seals to be in the corresponding code region as closure seals.
+That is because we cannot trust untrusted components to treat their seals correctly (e.g.
+not hand them out to other components, only sign return capabilities with return seals etc.), so instead, they must be considered as closure seals, which the component is essentially allowed to sign anything with.
 
@@
     \begin{array}{l}
+      \gc = (\ta,\stkb) \tand\\
       \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
@@
   \end{itemize}
+\end{lemma}
+
+\begin{lemma}[Adequacy of logical relation]
+  If
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$
+  \item $i \leq n$
+  \item $\Phi_S \sterm[i]{\gc}$
+  \end{itemize}
+  then $\Phi_T\term$.
+
+  Also, if
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq](W)$
+  \item $i \leq n$
+  \item $\Phi_T \term[i]$,
+  \end{itemize}
+  then $\Phi_S\sterm{\gc}$.
 \end{lemma}

### 2018-04-11 fully-abstract-comp/technical-report.tex
Fix problem in last commit and remove unnecessary requirement in readXCondition.
Resolve another TODO: an FTLR for components.

@@
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
-      \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{R(r),C(r),M(r),\gc}
+      % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
+      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
     \end{array}
@@
     \end{itemize}
+  \end{itemize}
+  and there exist $r_1 \cdots r_m \in \dom(\pwheap)$ such that
+  \begin{itemize}
+  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\pure}$
+  \item $\dom(\msdata) = \biguplus_i A_i$
   \end{itemize}
@@
 
+\subsection{FTLR for components}
+\label{sec:ftlr-for-comps}
+
+\begin{lemma}[FTLR for components]
+  If
+  \begin{itemize}
+  \item $\gc = (\ta,\stkb)$
+  \item One of the following holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\var{comp}$ is a reasonable component (see
+      Section~\ref{sec:full-abstraction}) and $\dom(\comp.\mscode) \subseteq \ta$
+    \item $\trust = \untrusted$ and $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
+    \end{itemize}
+  \item $W \vdash \var{comp}$
+  \end{itemize}
+  Then $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$.
+\end{lemma}
+Note: the trusted case of the above lemma can be considered as a compiler correctness result.
+The untrusted case can be considered as a back-translation correctness result.
+
 \subsection{Related execution configurations}
@@
 \section{Full Abstraction}
+\label{sec:full-abstraction}
 \lau{TODOs from last day in Leuven}
@@
 \lau{RESOLVED?: Component logical relation. Rough sketch is in white-board-notes/leuven-23-3-2018-3.jpg (top left).}
-\dominique{RESOLVED?: Compatibility lemma for linking: if $c_1$ and $c_1'$ are related and $c_2$ and $c_2'$, then so should be $c_1\bowtie c_2$ and $c_1' \bowtie c_2'$ (if defined).}
-\lau{TODO: Version of FTLR for programs.}
+\dominique{RESOLVED: Compatibility lemma for linking: if $c_1$ and $c_1'$ are related and $c_2$ and $c_2'$, then so should be $c_1\bowtie c_2$ and $c_1' \bowtie c_2'$ (if defined).}
+\lau{RESOLVED: Version of FTLR for programs.}
 \lau{RESOLVED: Compatibility lemma for initial executable configuration construct, i.e.

### 2018-04-11 fully-abstract-comp/technical-report.tex
Comment out obsolete example, remove old back-translation section.

@@
 
-\clearpage
-\section{Examples}
-\subsection{Capability Opacity}
-\label{subsec:capability-opacity}
-
-\dominique{27-09-2017: Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
-\lau{28-09-2017: It seems to me that this is indeed solved by the variable length instructions.
+% \clearpage
+% \section{Examples}
+% \subsection{Capability Opacity}
+% \label{subsec:capability-opacity}
+
+% \dominique{27-09-2017: Is this example still relevant? Isn't it solved differently by the variable-length-instructions idea?}
+% \lau{28-09-2017: It seems to me that this is indeed solved by the variable length instructions.
 That said, I would like to keep it around (maybe change it so it becomes an argument for variable length instructions).}
-This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{\offpc,\offsigma}{}{}$ and $\sreturn$ instructions.
+% This example was introduced when we envisioned a capability machine with $\spush{}$, $\spop{}$, $\ssload{}{}$, $\scall{\offpc,\offsigma}{}{}$ and $\sreturn$ instructions.
 The below example is a motivation for having variable length instructions in the source language because if we have enough memory to ``do the same'' in the target language as in the source language, then the below example does not work.
 
-The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-p1 ::= if r1 is length 2 then
-         call r1 with the following callback in r5:
-           {put $\textit{\texttt{diverging}}$ closure in r2;
-            return};
-         halt
-       else diverge
-
-p2 ::= if r1 is length 2 then
-         call r1 with the following callback in r5:
-           {put $\textit{\texttt{terminating}}$ closure in r2;
-            return};
-         halt
-       else diverge
-\end{lstlisting}
-The diverging closure could just contain $\sjmp{\pcreg}$, which diverges. The terminating closure could just be $\sreturn$.
-
-The context with $r1$ as an executable capability pointing to:
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-$\scall{\offpc,\offsigma}{r5}{0}$
-$\sjmp{r5}$
-\end{lstlisting}
-\footnote{At this time $\scall{\offpc,\offsigma}{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
+% The following pseudo program demonstrates the need of opaque capabilities. If we assume a system with no opaque capabilities, then the following programs break compiler correctness
+% \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+% p1 ::= if r1 is length 2 then
+%          call r1 with the following callback in r5:
+%            {put $\textit{\texttt{diverging}}$ closure in r2;
+%             return};
+%          halt
+%        else diverge
+
+% p2 ::= if r1 is length 2 then
+%          call r1 with the following callback in r5:
+%            {put $\textit{\texttt{terminating}}$ closure in r2;
+%             return};
+%          halt
+%        else diverge
+% \end{lstlisting}
+% The diverging closure could just contain $\sjmp{\pcreg}$, which diverges. The terminating closure could just be $\sreturn$.
+
+% The context with $r1$ as an executable capability pointing to:
+% \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+% $\scall{\offpc,\offsigma}{r5}{0}$
+% $\sjmp{r5}$
+% \end{lstlisting}
+% \footnote{At this time $\scall{\offpc,\offsigma}{}{}$ did argument spilling and it only takes one argument because we were considering enter capabilities.}distinguishes the two contexts, but two instructions are not enough to do the same at the target level.
 We would not have enough instructions to set up a proper return pointer for the compiled return to use.
-
-\clearpage
-\section{Back translation}
-The back translation is an embedding of the target language into the source language.
 

### 2018-04-11 fully-abstract-comp/technical-report.tex
Add a compatibility lemma for context plugging.
Uncomment resolved TODO.

@@
 
+\begin{lemma}[Compatibility lemma for context plugging]
+  If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
+  $\npair{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \oplus W_2)$.
+\end{lemma}
+
 \section{Full Abstraction}
@@
 
-\dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures).
-That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.}
+% \dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures).
+That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.
+Update: this is resolved}
 \begin{definition}

### 2018-04-11 fully-abstract-comp/ftlr-proof-cases.tex
Fix compilation of ftlr-proof-cases.tex
Update the proof of a lemma supporting the FTLR.

@@
 \[
-  \Phi_T \nstep[j] \Phi_T' \text{ and } \npair[n-j]{(\Phi_S,\Phi_T')} \in \lror,
+  \Phi_T \nstep[j]{} \Phi_T' \text{ and } \npair[n-j]{(\Phi_S,\Phi_T')} \in \lror,
 \]
@@
 \[
-  \Phi_S \nstep[j] \Phi_S' \text{ and } \npair[n-j]{(\Phi_S',\Phi_T)} \in \lrol,
+  \Phi_S \nstep[j]{\gc} \Phi_S' \text{ and } \npair[n-j]{(\Phi_S',\Phi_T)} \in \lrol,
 \]
@@
   \begin{itemize}
-  \item $\Phi_S \nstep[j_S] \Phi_S'$
-  \item $\Phi_T \nstep[j_T] \Phi_T'$
+  \item $\Phi_S \nstep[j_S]{\gc} \Phi_S'$
+  \item $\Phi_T \nstep[j_T]{} \Phi_T'$
   \item $\npair{(\Phi_S',\Phi_T')} \in \lro$
@@
 \begin{proof}
-  From $\npair{(b,e)}\in\readCond{l,W}$, we get an $r \in \addressable{\lin,
-    \pwheap}$ and $[\baddr',\eaddr'] \supseteq [\baddr,\eaddr]$ such that $\pwheap(r)
-  \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}$.
+  From $\npair{(b,e)}\in\readCond{l,W}$, we get an $S \subseteq \addressable{\lin,
+    \pwheap}$, an $R : S \rightarrow \powerset{\mathbb{N}}$ with $\bigcup_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwheap(r)
+  \nsubeq \stdreg{R(r)}{\pur}$ for all $r \in S$.
+
+  Since $a \in [b,e]$, there is a unique $r \in S$ such that $a \in R(r)$.
 
@@
   $\stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m)$,
-  $\ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m$,
+  $\ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m$ is defined,
   $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
-  $\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}$ such that
+  $\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}$, $\ms_{T,f}$, $\ms_{S,f}$, $\ms_S'$ such that
   \begin{itemize}
+  \item $\ms_S =\ms_{f,S} \uplus \ms_S'$
   \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus
-    \ms_\var{T,heap}$
+    \ms_\var{T,heap} \uplus \ms_{T,f}$
   \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
   \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
-  \item $\memSatHeap{\ms_S,\ms_\var{T,heap}}{W_\var{heap}}$.
+  \item $\memSatHeap{\ms_S',\ms_\var{T,heap}}{W_\var{heap}}$.
   \end{itemize}
@@
   $W_\var{heap} = \oplus_{r \in \dom(\activeReg{\pwheap[W_\var{heap}]})} R_W(r)$,
-  $\forall r \in \activeReg{\pwheap[W_\var{heap}]}, n' < n$, we have that
-  $\npair[n']{R_\ms(r)} \in  \pwheap[W_\var{heap}](r).H \; \pwheap[W_\var{heap}](r).s \; \xi^{-1}(R_W(r))$.
+  $\forall r \in \activeReg{\pwheap[W_\var{heap}]}$, we have that
+  $\npair{R_\ms(r)} \in  \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
 
   We have that $r \in \addressable{\lin, \pwheap} \subseteq
-  \activeReg{\pwheap[W_\var{heap}]}$, so $\npair[n']{R_\ms(r)} \in
-  \pwheap[W_\var{heap}](r).H \; \pwheap[W_\var{heap}](r).s \; \xi^{-1}(R_W(r))$.
-  Because $\pwheap(r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
-  also $\pwheap[W_\var{heap}](r) \nsubeq \stdreg{[\baddr',\eaddr']}{\pur}$ this
-  means that also $\npair[n']{R_\ms(r)} \in
-  H^{\mathrm{std}}_{[\baddr',\eaddr']}\; \xi^{-1}(R_W(r))$.
+  \activeReg{\pwheap[W_\var{heap}]}$, so $\npair{R_\ms(r)} \in
+  \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
+  Because $\pwheap(r) \nsubeq \stdreg{R(r)}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
+  also $\pwheap[W_\var{heap}](r) \nsubeq \stdreg{R(r)}{\pur}$ this
+  means that also $\npair{R_\ms(r)} \in
+  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined $\nsubeq$ somewhere?}
 
-  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = [\baddr',\eaddr']$ and we get a $S : [\baddr',\eaddr'] \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in [\baddr',\eaddr']} S(\aaddr)$ and $\forall \aaddr \in [\baddr',\eaddr'] \ldotp \npair[n'-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
-  \dominique{TODO: get rid of n'-1 above somehow.}
+  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r)$ and we get a $S : R(r) \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in R(r)} S(\aaddr)$ and $\forall \aaddr \in R(r),n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
 
-  Since $a \in [b,e] \subseteq [\baddr',\eaddr']$, we have that $\npair[n'-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
-  we can take $W_M' = W_r' \oplus W_{\var{heap}}' \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})$ with $W_r' = \oplus_{\aaddr \in ([\baddr',\eaddr']\setminus \{\aaddr\})} S(\aaddr)$ and $W_\var{heap}' = \oplus_{r' \in (\dom(\activeReg{\pwheap[W_\var{heap}]})\setminus \{r\})} R_W(r')$, and get
+  Since $\aaddr \in R(r)$ and $n' < n$, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
+  we can take $W_M' = W_r' \oplus W_{\var{heap}}' \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})$ with $W_r' = \oplus_{\aaddr \in (R(r)\setminus \{\aaddr\})} S(\aaddr)$ and $W_\var{heap}' = \oplus_{r' \in (\dom(\activeReg{\pwheap[W_\var{heap}]})\setminus \{r\})} R_W(r')$, and get
   \begin{align*}
@@
     &=
-    \oplus_{\aaddr \in [\baddr',\eaddr']} S(\aaddr) \oplus W_{\var{heap}}'\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
+    \oplus_{\aaddr \in R(r)} S(\aaddr) \oplus W_{\var{heap}}'\oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})\\
     &=
@@
 
-  Additionally, if $\npair{(b,e)}\in\writeCond{l,W}$, then
-  we get an $r' \in \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$
-  and $[\baddr',\eaddr'] \supseteq [\baddr,\eaddr]$ such that
-  $\pwheap(r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur}$ and
-  $\pwheap(r) \text{ is address-stratified}$. Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwheap[W_\var{heap}](r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur}$ and
+  Additionally, if $\npair{(b,e)}\in\writeCond{l,\pwheap}$, then
+  we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
+  for all $r \in S'$, $\pwheap(r) \nsupeq \stdreg{R'(r)}{\pur}$ and
+  $\pwheap(r) \text{ is address-stratified}$.
+
+  Since $a \in [b,e]$, there is an $r' \in S'$ such that $a \in R'(r')$ .
+
+  Because $W \oplus W_M$ is defined, it follows that also 
+  $\pwheap[W_\var{heap}](r) \nsupeq \stdreg{R(r)}{\pur}$ and
   $\pwheap[W_\var{heap}](r) \text{ is address-stratified}$.
-  \dominique{TODO: get rid of n-1 above.}
 
-  It follows that $r = r'$ because $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = [\baddr',\eaddr'] = \dom(R_\ms(r').2) = \dom(R_\ms(r').1)$ and all the $R_\ms(r).1$ and $R_\ms(r).2$ are disjoint.
+  It follows that $r = r'$ because $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r) = \dom(R_\ms(r').2) = \dom(R_\ms(r').1)$ and all the $R_\ms(r).1$ and $R_\ms(r).2$ are disjoint.
 
-  We have that $\npair[n']{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \;
-  \pwheap[W_\var{heap}](r).s \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupeq[n-1] \stdreg{[\baddr',\eaddr']}{\pur}$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupeq[n-1] \stdreg{R(r)}{\pur}$.
 

### 2018-04-13 fully-abstract-comp/ftlr-proof-cases.tex
Prove simple lemma.

@@
 
+\begin{lemma}[purePart is idempotent]
+  For all $W$, $\purePart(\purePart(W)) = \purePart(W)$
+  \label{lem:purePart-idempotent}
+\end{lemma}
+\begin{proof}
+  Follows easily from the definition.
+\end{proof}
+
 \begin{lemma}[non-linear words are pure]
@@
 \end{lemma}
+\begin{proof}
+  Follows easily by inspecting the definitions of $\lrv$, $\readCond$, $\addressable$, $\writeCond$, $\execCond$ and $\xReadCond$ and using Lemma~\ref{lem:purePart-idempotent}.
+\end{proof}
 

### 2018-04-13 presentations/esop18-04-2018/main.tex
added awkward example as first example (probably want to use a different one)
presentaion

--- 2017-12-19 presentations/illustrations/attacks.tex
@@
+\documentclass{beamer}
+\addtobeamertemplate{navigation symbols}{}{%
+    \usebeamerfont{footline}%
+    \usebeamercolor[fg]{footline}%
+    \hspace{1em}%
+    \insertframenumber/\inserttotalframenumber
+  }
+
+\usepackage{listings}
+\renewcommand{\ttdefault}{pcr}
+\lstset{
+  basicstyle=\ttfamily,
+  keywordstyle=\ttfamily\bfseries,
+  morekeywords={let, ref, assert, in},
+  columns=fullflexible,
+  keepspaces=true,
+  escapeinside={(*}{*)}
+}
+
+\input{../illustrations/preamble}
+\input{loc_preamble}
+\title{Reasoning About a Machine with Local Capabilities}
+\subtitle{Provably Safe Stack and Return Pointer Management}
+\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
+\institute{\inst{1} Aarhus University \and %
+  \inst{2} imec-DistriNet, KU Leuven}
+
+\date{ESOP, April 17, 2018}
+
+\tikzset{
+    gray on/.style={alt=#1{gray}{}},
+    alt/.code args={<#1>#2#3}{%
+      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
+    },
+}
+
+\begin{document}
+\maketitle
+
+\begin{frame}[fragile]
+  \frametitle{How Do We Reason About Programs Informally}
+  % Motivation
+  \centering
+
+  \begin{tabular}{c}
+    \begin{lstlisting}
+let x = ref 0 in
+  (*$\lambda$*)f.(x := 0;
+      f();
+      x := 1;
+      f();
+      assert(x == 1))
+    \end{lstlisting}
+  \end{tabular}
+
+  \note{Informally, what does this program do?}
+  \note{It allocates a reference to 0 and saves a closure to x.
+The closure takes a callback, assigns 0 to x, calls the callback, assigns 1 to x, calls the callback another time and finally asserts x to be 1.}
+  \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
+  \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point.
+The low-level machine needs to enforce this if I want to be able to do have this closure interact with arbitrary machine code.}
+\end{frame}
+
 \begin{frame}
-  \frametitle{Stack and return capabilities: Attack 1}
+  \frametitle{Traditional Stack Pointers}
+  \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
+    \end{scope}
+    \draw (-1.5,0) node {};
+    \draw (-1.5,15) node {};
+    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
+
+    % traditional:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
+
+    % traditional call:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
+
+    % traditional return:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
+
+    % traditional call, attack:
+    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
+    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
+    \draw<.-> (8,7) node[draw=black] (sp0) {};
+    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
+
+    % read other stack frames
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
+    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
+
+    % break well-bracketedness
+    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
+    \draw<.-> (8,7) node[draw=black] (sp0) {};
+    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
+    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
+      skip caller frame};
+  \end{tikzpicture}
+  \note{We need something to enforce security properties on a low-level machine.}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Capability Machine}
+  % short description of capability machines
+
+  \begin{columns}
+\begin{column}{0.5\textwidth}
+  \begin{itemize}
+  \item Low-level machine
+    \note{Has instructions as you would expect, load, store, jmp, etc.}
+  \item Capabilities replace pointers
+    \begin{itemize}
+    \item Pointer
+    \item Range of authority
+    \item Kind of authority
+      \begin{itemize}
+      \item read/write/execute
+      \item enter
+      \end{itemize}
+      \note{Has instructions for manipulating capabilities.}
+    \end{itemize}
+    \item Authority checked dynamically
+  \end{itemize}
+\end{column}
+\begin{column}{0.5\textwidth}
+    \begin{center}
+      % Insert capability figure
+     \end{center}
+\end{column}
+\end{columns}
+  \note{This particular setup gives a very fine-grained memory control.}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack and Return Capabilities: Attack 1}
+  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
 
+    \note{Evil callee stores stack pointer on the heap.}
     % stores stack pointer in heap
@@
 
+    \note{Evil callee returns to the caller.}
     % attacker returns
@@
 
+    \note{caller uses a bit more of the stack and calls the evil callee again.}
+    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
     % attacker gets called again + uses old stack pointer.
@@
 \begin{frame}
-  \frametitle{Stack and return capabilities: Attack 2}
+  \frametitle{Local Capabilities}
+  % Short description of local capabilities
+  \begin{itemize}
+  \item Capabilities tagged with locality (local or global)   \note{We call non-local capabilities global}
+  \item New \emph{write-local} permission.
+  \item Local capabilities can only be stored by capabilities with \emph{write-local} permission
+  \end{itemize}
+  Calling convention highlights
+  \begin{itemize}
+  \item Stack capability is local with permission read, write-local, and execute.
+  \item Clear stack before passing stack capability to untrusted code.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Local Stack Capabilities Prevent Attack 1}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker...
+    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
+    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    \draw node[right=0cm of sp1] { stack pointer };
+
+    % stores stack pointer in heap
+    \begin{scope}
+      \fill[gray!20] (14,6) rectangle (18,13);
+      \draw (14,6) -- (14,13);
+      \draw (18,6) -- (18,13);
+    \end{scope}
+    \draw (16,14) node {heap memory};
+    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
+    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
+    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
+    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack and Return Capabilities: Attack 2}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
   \end{tikzpicture}
+  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
 \end{frame}
@@
 \begin{frame}
+  \frametitle{Calling Convention (Continued)}
+  \begin{itemize}
+  \item $\dots$
+  \item Clear stack and non-argument registers before invoking untrusted code.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stack Clearing Prevents Attack 2}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    \frametitle{Stack and return capabilities: Attack 3}
-    \begin{scope}
-    \clip (-.1,-.1) rectangle (5.1,10.1);
-    \fill[fill=gray!50] (0,0) rectangle (5,10);
-    \draw (0,0) -- (0,10);
-    \draw (5,0) -- (5,10);
-    \end{scope}
-    \draw (2.5,11) node {code memory};
+    % recurrent parts
+    \stdstackstart
+    % attacker stores copy of stack pointer high in the stack
+    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \begin{onlyenv}<+>
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
+    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
+    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);
 
-    \begin{scope}
-    \clip (12.9,3.9) rectangle (18.1,11.1);
-    \fill[fill=gray!50] (13,4) rectangle (18,11);
-    \draw (13,4) -- (13,11);
-    \draw (18,4) -- (18,11);
-    \end{scope}
-    \draw (15.5,12) node {more code memory};
-
-    \begin{scope}
-    \clip (12.9,-.1) rectangle (18.1,2.1);
-    \fill[fill=gray!50] (13,0) rectangle (18,2);
-    \draw (13,0) -- (13,2);
-    \draw (18,0) -- (18,2);
-    \end{scope}
-    \draw (15.5,3) node {heap memory};
-
-\note{Caller calls callee}
-    \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-
+    % attacker calls trusted code again
     \begin{onlyenv}<+->
-      \capbrace[ret1]{(5.1,1)}{(5.1,4)}
-      \draw node[fill=gray!50, right=0cm of ret1] { \phantom{ret-ptr} };
-      \draw node[right=0cm of ret1] { ret-ptr };
+      \inactadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
     \end{onlyenv}
 
-\note{Callee has access to some place to store things.}
+    % trusted code clears the stack
     \begin{onlyenv}<+->
-      \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=white] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
+      \foreach \x in {9,9.5,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
+    (13,13.5) -- (10,14.5);
+
+    % attacker uses old stack pointer
+    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \begin{onlyenv}<.>
+      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
     \end{onlyenv}
 
-    \begin{onlyenv}<.-.(2)>
-      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
-      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
-      \draw (18,5.5)  edge[->,bend left] (copy);
+    % attacker returns
+    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
+    \begin{onlyenv}<.>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
     \end{onlyenv}
 
-\note{Callee stores the return pointer they got from caller.}
+    % trusted code clears stack again
     \begin{onlyenv}<+->
-      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-      \draw (13,1.25) edge[->,red,thick,bend left] (ret1);
+      \foreach \x in {7,7.5,...,14}
+      {
+        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+    \end{onlyenv}
+    % then, the trusted code can return
+    \begin{onlyenv}<+>
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
     \end{onlyenv}
 
-\note{Callee calls callee (1)}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=3pt,raise=4pt},yshift=0pt]
-      (12.9,5) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
-      \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
-      \draw node[left=0cm of ret2] (retptr1) { ret-ptr (1) };
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{(Full) Calling Convention}
+  % and stack management
+  \note{We need to do a few more things, but this is the rest (just to illustrate that it is fairly simple).}
+  \note{Every point is motivated by some attack.}
+  \begin{itemize}
+  \item Initially:
+    \begin{itemize}
+    \item Stack capabilitiy local capability with read, write-local, and execute authority.
+    \item No global write-local capabilities on the machine.
+    \end{itemize}
+  \item Prior to returning to untrusted code:
+    \begin{itemize}
+    \item Clear the stack.
+    \item Clear non-return registers.
+    \end{itemize}
+  \item Prior to calls to untrusted code:
+    \begin{itemize}
+    \item Push activation record to the stack and create enter-capability.
+    \item Restrict the stack pointer to the unused part and clear that part.
+    \item Clear non-argument registers.
+    \end{itemize}
+  \item Only invoke global call-backs.
+  \item When invoked by untrusted code
+    \begin{itemize}
+    \item Make sure the stack pointer has read, write-local and execute authority.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Formalizing the Guarantees of a Capability Machine}
+  \note{How can we be sure calling convention works.
+Specifically, if a program interacts with intrusted code using the CC, can we formally show the correctness of the program if it relies on well-bracketedness or local-state encapsulation.}
+
+  \note{Need formal statement of the guarantees provided by the capability machine including the specific guarantees for local capabilities.}
+  \begin{itemize}
+  \item How do we know the calling convention works?
+  \item Unary step-indexed Kripke logical relation over recursive worlds
+    \begin{itemize}
+    \item Statement of guarantees probided by the capability machine
+    \end{itemize}
+  \end{itemize}
+  \note{We state this formal statement in terms of a unary step-indexed Kripke logical relation over recursive worlds}
+
+  \note{Calling convention main application, but it is very general - can be used to reason about other programs.}
+
+  \note{In the following: give some intuition for different parts of LR}
+
+  \note{correctness here could be assert not violated}
+
+  \note{mention better than previous. Define the guarantees}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Worlds, Safe Values, and Step-Indexing}
+\begin{columns}
+  \begin{column}{0.5\textwidth}
+    \begin{itemize}
+    \item<1-> Capabilities represent bound on executing code
+    \item<2-> World, $W$      
+      \begin{itemize}
+      \item Collection of invariants
+      \end{itemize}
+    \item<4-> Predicate for safe values w.r.t world, $\stdvr(W)$
+      \begin{itemize}[<8->]
+      \item Recursively defined
+      \end{itemize}
+    \end{itemize}
+\end{column}
+\begin{column}{0.5\textwidth}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \scope
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \fill[fill=white] (0,0) rectangle (6,15);
+      \draw (0,0) -- (0,15);
+      \draw (6,0) -- (6,15);
+    \endscope
+    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \draw (-1.5,0) node {};
+    \draw (-1.5,15) node {};
+    \foreach \x in {0.5,1,...,14}
+    {
+      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
+    };
+
+    % One capability
+    \draw<1-6> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
+    (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
+    \draw<1-4>  (6,7.25) edge[black,->,in=45] (sp1);
+
+    % Invariant colored
+    \fill<2->[red,opacity=.7] (0,1) rectangle (6,6);
+    \fill<2->[blue,opacity=.7] (0,7) rectangle (6,10);
+
+    \fill<3-6>[black, fill opacity=0, text opacity=0.7] (0,4) rectangle (6,4.5) node[pos=.5] {\footnotesize 42};
+
+    \draw<5> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{r}} (sp1)   ;
+    \draw<6> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{rw}} (sp1)   ;
+    \draw<6>[very thick,color=red] (6.5,6.5) -- (8.5,4.5)
+    (6.5,4.5) -- (8.5,6.5);
+
+    \draw<7-> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
+    (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
+    \path<7->  (6,4.25) edge[black,out=30,in=135] node[above right]  {\footnotesize \textsc{r}} (8,5) 
+              (8,5) edge[black,->,out=-45,in=45](sp1);
+
+  \end{tikzpicture}
+
+\end{column}
+\end{columns}
+  % Do this as a series of figures?
+
+  % Figure with memory and one capability to make the bound explicit
+  \note{compared to normal assembly, capabilities represent bounds on executing block of code.}
+
+  \note{we have no observable I/O, so the authority bounds we consider are related to memory. However, more fine-grained/detailed than read/write-authority.}
+
+  % Color the memory to illustrate different regions of the world
+  \note{a piece of code can be bound by arbitrary memory invariants which we define in a \emph{world}.}
+
+  % Write down a static invariant (some address contains 42, maybe write a legend on the left side of slide.).
+  \note{essentially, a world is a collection of invariants and safety of words defined with respect to a world.}
+
+  \note{Define a set of words that are safe w.r.t. world W V(W) in P(Word)}
+
+  \note{Whether a capability is safe depends on the authority it carries}
+
+  % Add a read capability to the address that contains 42, safe
+  % Add read/write capability to the address that contains 42, not safe - can be used to violate the invariant.
+  \note{Example, w.r.t. a world with the invariant that an address contains constant. Safe for read capability, not write as write capability can break this invariant.}
+
+  % Add a read capability that refers to itself
+  \note{safety for a read capability is the case if the read capability only gives access to safe capabilities.}
+
+  \note{What if that part of memory contains a read capability for the same part of memory? Cyclic definition. Solved by step-indexing.}
+
+  \note{related to similar issue with languages with recursive types or higher-order ML-list references.}
+
+  \note{Solved by step-indexing - safety up to a certain number of interaction steps.}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Future Worlds and Invariants, and Recursive Worlds}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \scope
+      \clip (-.1,-.1) rectangle (6.1,8.1);
+      \fill[fill=white] (0,0) rectangle (6,8);
+      \draw (0,0) -- (0,8);
+      \draw (6,0) -- (6,8);
+    \endscope
+    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \draw (-1.5,0) node {};
+    \draw (-1.5,8) node {};
+    \foreach \x in {0.5,1,...,7}
+    {
+      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
+    };
+
+    \fill[red,opacity=.7] (0,7.5) rectangle (6,6);
+    \fill[blue,opacity=.7] (0,6) rectangle (6,4);
+
+    \begin{onlyenv}<2->
+    \scope
+      \clip (11.9,-.1) rectangle (18.1,8.1);
+      \fill[fill=white] (12,0) rectangle (18,8);
+      \draw (12,0) -- (12,8);
+      \draw (18,0) -- (18,8);
+    \endscope
+    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \draw (-1.5,0) node {};
+    \draw (-1.5,8) node {};
+    \foreach \x in {0.5,1,...,7}
+    {
+      \draw[fill=white] (12,\x) rectangle (18,\x+.5) node[pos=.5,color=teal] {};
+    };
+
+    \fill[red,opacity=.7] (12,7.5) rectangle (18,6);
+    \fill[blue,opacity=.7] (12,6) rectangle (18,4);
+
+    \draw [decorate,decoration={brace,amplitude=3pt,raise=2pt},yshift=0pt]
+    (11.75,1) -- (11.75,4) node[draw=black] (sp1) [black,midway,xshift=-12pt] {}; % ,shape=coordinate,xshift=1pt
+    \draw  (10,0.5) edge[black,->,in=225] (sp1);
+
+    \draw (7,7.5) edge [black,thick, ->] node[above] {\footnotesize Memory allocated} (11,7.5);
   \end{onlyenv}
+  \begin{onlyenv}<3->
+    \fill[green,opacity=.7] (12,4) rectangle (18,1);
 
-    \begin{onlyenv}<+->
-      \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=gray!30] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
+    \draw (9,4) node {\huge $\sqsubseteq$};
   \end{onlyenv}
+    %Invariant transition systems
+    \begin{onlyenv}<4->
+    \node[circle,draw=black, fill=white] (a) at (3,6.75) {};
+    \draw[>=stealth] (2,6.75) edge[black,->] (a);
 
-\note{Callee (1) calls callee (2)}
-    \begin{onlyenv}<+->
-      \capbrace[ret3]{(5.1,5)}{(5.1,8)}
-      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-      \draw node[right=0cm of ret3] (retptr2) { ret-ptr (2) };
-    \end{onlyenv}
+    \node[circle,draw=black, fill=white] (b) at (2,5) {};
+    \draw[>=stealth] (1,5) edge[black,->] (b);
+    \node[circle,draw=black, fill=white] (c) at (4,5) {};
+    \draw[>=stealth] (b) edge[black,->, bend left] (c);
+    \draw[>=stealth] (c) edge[black,->, bend left] (b);
 
-    \begin{onlyenv}<+->
-      \draw[fill=gray!30] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
-      \draw[fill=white] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
-    \end{onlyenv}
-\note{Callee (2) colludes with callee}
-    \begin{onlyenv}<+->
-      \draw (21,7.5) node[devil,opaque=0.5,mirrored,minimum size=1cm] {};
-    \end{onlyenv}
-\note{Callee (2) has access to ret-ptr (the return pointer caller passed to callee)}
-    \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
-      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
-      \draw (18,8.25)  edge[->,red,thick,bend left] (copy);
-      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-    \end{onlyenv}
+    \node[circle,draw=black, fill=white] (a2) at (15,6.75) {};
+    \draw[>=stealth] (14,6.75) edge[black,->] (a2);
 
-\note{Callee (2) uses ret-ptr to return}
-    \begin{onlyenv}<+->
-      \draw[fill=gray!30] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
-      \draw[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
-      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
-    \end{onlyenv}
+    \node[circle,draw=black, fill=white] (b2) at (14,5) {};
+    \draw[>=stealth] (13,5) edge[black,->] (b2);
+    \node[circle,draw=black, fill=white] (c2) at (16,5) {};
+    \draw[>=stealth] (b2) edge[black,->, bend left] (c2);
+    \draw[>=stealth] (c2) edge[black,->, bend left] (b2);
 
-\note{This breaks well-bracketedness as ret-ptr (2) and ret-ptr (1) was not returned to}
-    \begin{onlyenv}<+->
-      \draw node [ellipse, fit=(retptr1) (retptr2), draw, red] { };
-      \draw node[red,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  
+    \node[circle,draw=black, fill=white] (d) at (14,2.5) {};
+    \draw[>=stealth] (13,2.5) edge[black,->] (d);
+    \node[circle,draw=black, fill=white] (e) at (16,3.25) {};
+    \draw[>=stealth] (d) edge[black,->, bend left] (e);
+    \draw[>=stealth] (e) edge[black,->, bend left] (d);
+    \node[circle,draw=black, fill=white] (f) at (16,1.75) {};
+    \draw[>=stealth] (e) edge[black,->, bend left] (f);
   \end{onlyenv}
 \end{tikzpicture}
+
+  \begin{overprint}
+\onslide<1-4>
+  \begin{itemize}
+  \item<1-> Memory evolves over time
+  \item<3-> Add invariants in future worlds
+  \item<4-> Invariants as state machines
+  \end{itemize}
+  \onslide<5-6>
+  \begin{itemize}
+  \item<5-> Each state contains a predicate of accepted memory segments
+  \item<6-> World indexed
+  \end{itemize}
+  \[
+    H \; :  \only<5>{\phantom{\; \Worlds \fun {}}}\only<6>{\; \Worlds \fun} \mathrm{Pred}(\mathrm{MemSeg})
+  \]
+\end{overprint}
+  \note{Memory changes over time, for instance new memory may be allocated.}
+  \note{Allow worlds to evolve. New invariants can be added to handle freshly allocated memory.}
+  \note{Safety of words monotone w.r.t. worlds (which makes it into a Kripke Logical relation).}
 \end{frame}
-%%% Local Variables:
-%%% TeX-master: "presentation-illustrations"
-%%% End:
+
+% \begin{frame}
+%   \frametitle{Invariants and Recursive Worlds}
+%   \note{The purpose of a piece of memory may change over time. Invariants need to be more like protocols.}
+%   \note{Take a closer look at invariants.}
+%   \note{Each invariant is a state machine to model changes of memory invariants for a piece of code over time.}
+
+%   \note{An invariant contains a mapping from states to predicate of memsegs...
+
+%  for pieces of memory it accepts.
+Predicate may be world dependent (as the validity may depend on the remainder of the memory, which for instance is the case if it contains a capability(?)) or we might just want to say a specific address contains some safe value, but safety is world dependent.
+This means that the predicate must be world indexed.}
+% \end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Local Capabilities}
+  \texttt{f} is unknown code and \texttt{c} is a capability.
+  \begin{lstlisting}
+    f(c);
+    f(1)
+  \end{lstlisting}
+  \begin{itemize}
+  \item<2-> \texttt{c} global $\Rightarrow$ available in second invocation of \texttt{f}
+
+    \note{when c global, second invocation in a world where c is safe}
+  \item<3-> \texttt{c} local $\Rightarrow$ not available in second invocation of \texttt{f}
+    \note{when c local, second invocation in a world where c is not (necessarily) safe}
+    \note{c essentially revoked, so the invariants it relies on need not hold.}
+  \end{itemize}
+
+  \note{Two future world relations, one for all capabilities and one for non-local capabilities}
+  \note{If c global, then second invocation must happen in public future world, so c valid.}
+  \note{If c local, then second invocation may happen in a private future world.}
+  \begin{onlyenv}<4->
+  \begin{lemma}[Double monotonicity of value relation]
+    \begin{itemize}
+    \item If $\npair{w} \in \stdvr(W)$ and $W' \futurewk W$ then $\npair{w} \in
+      \stdvr(W')$.
+    \item If $\npair{w} \in \stdvr(W)$ and $W' \futurestr W$ and $w$ is
+      not a local capability, then $\npair{w} \in \stdvr(W')$.
+    \end{itemize}
+  \end{lemma}
+\end{onlyenv}
+  \note{How does local/global capabilities affect all this.}
+  \note{If we hand a global capability to untrusted code, then it may be stored in memory, so we will only be able to reinvoke that code if we can guarantee that those values are still valid. }
+  \note{Formally, the worlds contain the invariants that the global capability depend on and reinvocation is only possible in future worlds where these invariants are respected.}
+  \note{Local capabilities on the other provides a means to revoke capabilities.
+If we invoke untrusted code and give them a local capability, then they have no way to store it aside from the register file (and the stack), so when they return we can be sure that the local capability resides nowhere but the register file and the stack.}
+  \note{If we clear the register file and stack for old local capabilities before we invoke the untrusted code again, then we have essentially revoked those local capabilities and we can safely invoke the untrusted code again with the assumption that it does not have access to the local capability from the first call.}
+  \note{Formally, we allow this call to happen in a \emph{private future} world in which all global capabilities remain safe, but local capabilities don't.}
+  \note{We also have a \emph{public future} world relation in which \emph{all} capabilities remain safe.}
+  \note{Concretely, worlds may contain temporary regions on which local capabilities may depend.
+These regions persist in public future worlds which allows the local capabilities to rely on them.
+On the other hand, in private future worlds temporary regions may be revoked which is why local capabilities may not be safe in private future worlds.}
+
+  \note{include comment about priv/pub future worlds + interplay with monotonicity?}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Fundamental Theorem of Logical Relations}
+    \begin{itemize}
+  \item General statement about the guarantees provided by the capability machine.
+  \item Intuitively: any program is safe as long as it only has access to safe values.
+  \end{itemize}
+  \begin{theorem}[Fundamental theorem (simplified)]
+      If 
+      \[
+        \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+      \]
+      then
+      \[
+        \npair{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W)
+      \]
+  \end{theorem}
+
+  % Comments/present high-level perspective
+  \note{readCond is the assumption that every thing in the interval [a,e] is safe to read.}
+  \note{$\stder$ is safe to execute relation. That is, it will respect all the memory invariants.}
+  \note{That is take an arbitrary capability. If it only has access to safe capabilities then it will preserve the invariants of the world.}
+  \note{Remeber, dynamic checks = failing is concidered secure} 
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{``Awkward Example''}
+    \begin{tabular}{c}
+    \begin{lstlisting}
+let x = ref 0 in
+  (*$\lambda$*)f.(x := 0;
+      f();
+      x := 1;
+      f();
+      assert(x == 1))
+    \end{lstlisting}
+  \end{tabular}
+\note{example known from the litterature}
+\note{even just in ML difficult as f can be the closure.}
+\note{the assert can fail if the calls are not well-bracketed!}  
+\note{the local state is difficult to handle as the closure and the contex needs to be able to update the invariant for x in different ways.
+(closure can switch between 0 and 1 as it pleases, but context can transition only from 0 to 1.)}
+\note{relies heavily on well-bracketedness}
+\note{we have made a faithfull translation and proved correctness (i.e., the assertion never fails).}
+\note{more semantic statement of guarantees allows us to do this.}
+\end{frame}
+
+% \begin{frame}
+%   \frametitle{Contributions}
+%   \begin{itemize}
+%     \item Formalisation of a capability machine with local capabilities
+%     \item Novel calling convention for enforcing control-flow correctness and local-state encapsulation
+%     \item A calling convention utilizing local capabilities to ensure well-bracketed control-flow and local-state encapsulation
+%     \item Step-indexed Kripke Logical Relation to reason about programs on said machine
+%       \begin{itemize}
+%         \item Single orthogonal closure
+%         \item Variant of public/private future world relation to express nature of local capabilities
+%       \end{itemize}
+%       %new insights for low-level systems?
+%     \item FTLR
+%       \begin{itemize}
+%       \item Semantic statement of guarantees offered by capability machine
+%       \end{itemize}
+%     \item Logical relation applied to a number of examples
+%       \begin{itemize}
+%       \item ``Awkward example''
+%       \end{itemize}
+%     \end{itemize}
+% \end{frame}
+
+\begin{frame}
+  \frametitle{Conclusion}
+  \begin{itemize}
+  \item Capability machines can guarantee properties of high-level languages
+  \item Calling convention for well-bracketedness and local-state encapsulation
+  \item Unary step-indexed Kripke logical relation over recursive worlds
+    \begin{itemize}
+    \item Formal statement about guarantees provided by capability machine
+    \item Reasoning about programs in general
+    \end{itemize}
+  \item Applied on the ``awkward example''
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \centering
+  Thank you!
+\end{frame}
+
+\end{document}

### 2018-04-13 fully-abstract-comp/ftlr-proof-cases.tex
Fix latex compilation.

@@
   \item If $\npair{(b,e)}\readCond{\lin}{W}$, then $\npair[n']{(b,e)}\readCond{\lin}{W}$.
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
   \end{itemize}
@@
 \begin{lemma}[purePart is idempotent]
-  For all $W$, $\purePart(\purePart(W)) = \purePart(W)$
+  For all $W$, $\mathit{purePart}( \purePart{W} ) = \purePart{W}$
   \label{lem:purePart-idempotent}
@@
   \item $w_S = \linCons{\ms_S(\aaddr)}$, $w_T = \linCons{\ms_T(\aaddr')}$
-  \item $\linConsPerm{\perm}{\ms_S(\aaddr)}$, $\linConsPerm{\perm}{\ms_T(\aaddr')}$
+  \item $\linConsPerm{\perm}{\ms_S(\aaddr)}$, $\linConsPerm{\perm'}{\ms_T(\aaddr')}$
+  \item $\aaddr \in [\baddr,\eaddr]$
+  \item $\aaddr' \in [\baddr',\eaddr']$
   \end{itemize}
@@
   \item $W_M = W' \oplus W_M'$
-  \item $\memSat{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$
-  \item $\npair{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$
+  \item $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$
+  \item $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$
   \end{itemize}
-  \dominique{11-3-2018: this lemma depends on Lemma~\ref{lem:non-linear-pure} for the non-linear case.}
 \end{lemma}
 \begin{proof}
-  TODO
+  From $\npair{(c,c')}\in\lrv(H_\sigma,W)$ with $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$, $\perm \in \readAllowed{}$ and $\perm' \in \readAllowed{}$, we get that $\baddr = \baddr'$, $\eaddr = \eaddr'$ and $\aaddr = \aaddr'$ and $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$.
+
+  Lemma~\ref{lem:readcond-writecond-work} then gives us a $W'$ and $W_M'$ such that $W_M = W' \oplus W_M'$ and $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$.
+
+  It remains to prove that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+  We have to distinguish the case that $\isLinear{\ms_S(\aaddr)}$ and the opposite case.
+  \begin{itemize}
+  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
+    From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
+    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+
+  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
+    The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by non-expansiveness of memory satisfaction, i.e.
+Lemma~\ref{lem:non-expansive}.
+  \end{itemize}
 \end{proof}

### 2018-04-13 fully-abstract-comp/technical-report.tex
Update the proof of a lemma supporting the FTLR.

@@
 \begin{definition}[The pure part of a world]
+  \label{def:purePart}
   For any world $W$, we define
@@
       \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
-      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
+      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \text{ is defined} \wedge\\
       W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
@@
       \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
+      \quad (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
       \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
@@
       \exists S : A \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in A \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \quad \forall \aaddr \in A, n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}

### 2018-04-15 presentations/esop18-04-2018/main.tex
Fix typo.
small changes in presentation
Presentation notes

@@
   }
+
+\usepackage{pgfpages}
+\setbeameroption{show notes}
+\setbeameroption{show notes on second screen=right}
 
@@
 \begin{frame}[fragile]
-  \frametitle{How Do We Reason About Programs Informally}
+  \frametitle{What Does This Program Do?}
   % Motivation
@@
       f();
-      assert(x == 1))
+      assert(!x == 1))
     \end{lstlisting}
@@
 
-  \note{Informally, what does this program do?}
-  \note{It allocates a reference to 0 and saves a closure to x.
+  \note<1>[item]{Consider program. Assuming a standard ML semantics we can say what it does.}
+  \note<1>[item]{Bind x to freshly allocated reference in a closure that...}
+  \note<1>[item]{takes callback f, sets x to 0, calls f, sets x to 1, calls f and finally asserts x points to 1.}
+  \note<1>[item]{Note the assumption that when we call f, then we return to a specific program point. This is what we call well-bracketedness and we assume we have this in many programming languages.}
+  \note<1>[item]{However, in order to execute this code, we need to compile it to assembly. }
+  \note<1>[item]{How is well-bracketedness guaranteed? In particular, how is it guaranteed if f is a piece of code we do not trust (maybe handwritten assembly).}
+  \note<2>[item]{\emph{We present a calling convention for capability machines that provide well-bracketedness and local state encapsulation as well as a logical relation that allows us to reason about such programs.}}
+  \note<2>[item]{Let's first consider how stack pointers traditionally are handled.}
+
+  % \note{Informally, what does this program do?}
+  % \note{It allocates a reference to 0 and saves a closure to x.
 The closure takes a callback, assigns 0 to x, calls the callback, assigns 1 to x, calls the callback another time and finally asserts x to be 1.}
-  \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
-  \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point.
+  % \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
+  % \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point.
 The low-level machine needs to enforce this if I want to be able to do have this closure interact with arbitrary machine code.}
@@
   \frametitle{Traditional Stack Pointers}
-  \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
+  % \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
+  \note<1-3>[item]{Simply put, a caller calls a function which}
+  \note<1-3>[item]{pushes a new stack frame on the stack the callee uses for its execution. }
+  \note<1-3>[item]{When the callee is done, then it returns to the caller by popping its stack frams.}
+  \note<4->[item]{If \emph{callee (evil) assembly code with no intention to follow the CC}, then there are multiple ways for them to break things:}
+  \note<5->[item]{Read or write directly from or to the caller's stack frame, breaking local-state encapsulation}
+  \note<6->[item]{Skip the caller's stack frame and return to one further down breaking well-bracketedness.}
+  \note<6->[item]{Clearly we need some kind of low-level enforcement mechanism.}
+
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
         % read other stack frames
-    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
     \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \draw<.-.(1)> (sp0) edge[->,out=235,in=0] (6,8.25);
+
+    % read other stack frames
+
+    \actadv[+]{(0,5.5)}{(6,8.5)}{callee stack frame}
     \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
@@
   \end{tikzpicture}
-  \note{We need something to enforce security properties on a low-level machine.}
+  % \note{We need something to enforce security properties on a low-level machine.}
 \end{frame}
@@
   \begin{itemize}
-  \item Low-level machine
-    \note{Has instructions as you would expect, load, store, jmp, etc.}
-  \item Capabilities replace pointers
+  \item<1-> Low-level machine
+    \note<1->[item]{Capability machines are low-level machines proposed in the systems community.}
+    \note<1->[item]{For instance, the CHERI OS operates on one.}
+    \note<1->[item]{Has all the instructions we expect, load, store, jmp, etc.}
+    % \note{Has instructions as you would expect, load, store, jmp, etc.}
+  \item<2-> Capabilities replace pointers
     \begin{itemize}
-    \item Pointer
-    \item Range of authority
-    \item Kind of authority
+    \item<3-> Pointer
+    \item<4-> Range of authority
+    \item<5-> Kind of authority
+      \note<5->[item]{Roughly two kinds of capabilities:}
       \begin{itemize}
-      \item read/write/execute
+      \item read, write, and execute
+        \note<5->[item]{Memory capabilities, allows you to do all the standard memory operations.}
       \item enter
+        \note<5->[item]{Provides encapsulation mechanism which allows seperation of security domains.}
+        \note<5->[item]{Can not be used for anything but jump, when jumped to becomes read/execute.}
       \end{itemize}
-      \note{Has instructions for manipulating capabilities.}
     \end{itemize}
-    \item Authority checked dynamically
+    \item<6-> Capability manipulation instructions
+    \item<7-> Authority checked dynamically
     \end{itemize}
 \end{column}
+
 \begin{column}{0.5\textwidth}
-    \begin{center}
-      % Insert capability figure
-     \end{center}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \draw (1,9) node {Memory};
+    \draw (-3.5,0) node {\phantom{.}};
+    \scope
+      \clip (-.1,-.1) rectangle (6.1,8.1);
+      \fill[fill=white] (0,0) rectangle (6,8);
+      \draw (0,0) -- (0,8);
+      \draw (6,0) -- (6,8);
+    \endscope
+    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \draw (-1.5,0) node {};
+    \draw (-1.5,8) node {};
+    \foreach \x in {0.5,1,...,7}
+    {
+      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
+    };
+
+    \draw<3>  (-3,1.5) edge[black,->,in=180] (0,2.75);
+
+    \draw<4-> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
+    (-0.25,1) -- (-0.25,4.5) node[draw=black] (sp1) [black,midway,xshift=-12pt] {};
+    \draw<4>  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize\phantom{\textsc{rw}}} (sp1);
+    \draw<5->  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize \textsc{rw}} (sp1);
+  \end{tikzpicture}
 \end{column}
 \end{columns}
-  \note{This particular setup gives a very fine-grained memory control.}
+ % \note{This particular setup gives a very fine-grained memory control.}
 \end{frame}
@@
   \frametitle{Stack and Return Capabilities: Attack 1}
-  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
+%  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
+  \note[item]<1->{Let's see how this changes things: Now the \emph{untrusted code cannot immediately read from the caller stack frame}, because the stack capability does not have authority over that part of memory.}
+  \note[item]<1->{There is nothing that prevents the \emph{untrusted code from storing the stk ptr on the heap.}}
+  \note[item]<3->{Upon return the \emph{callee regains its stack capability} which has authority over the callee stack frame and everything above. }
+  \note[item]<3->{The \emph{caller pushes some important things} on the stack and \emph{calls the untrusted code again}. With a smaller stack pointer.}
+  \note[item]<4->{The stack pointer the caller gives the untrusted code cannot be used to access the callee stack frame, but because the untrusted code stored the old stack pointer, it now has access to part of the callee's stack frame.}
+  \note[item]<4->{Again breaking local state encapsulation.}
+  \note[item]<4->{Need a way to make sure stack pointer is not stored for later use.}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
       (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,14) node[above right=1cm of sp1, xshift=-0.5cm] (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
     \end{onlyenv}
 
-    \note{Evil callee stores stack pointer on the heap.}
+%    \note{Evil callee stores stack pointer on the heap.}
     % stores stack pointer in heap
@@
     \end{onlyenv}
-    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \draw<.> (14,8.25) edge[->,red,thick,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
     \begin{onlyenv}<+->
@@
       (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
+      \draw (14,8.25) edge[->,red,thick,bend left] node[left] {\footnotesize \textsc{rw}}(spold);
     \end{onlyenv}
 
-    \note{Evil callee returns to the caller.}
+%    \note{Evil callee returns to the caller.}
     % attacker returns
@@
       (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,14) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[below right] {\footnotesize \textsc{rw}} (sp1);
       \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
@@
 
-    \note{caller uses a bit more of the stack and calls the evil callee again.}
-    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
+%    \note{caller uses a bit more of the stack and calls the evil callee again.}
+%    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
     % attacker gets called again + uses old stack pointer.
@@
         (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,14) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[below] {\footnotesize \textsc{rw}} (sp1);
       % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
   % Short description of local capabilities
+  CHERI inspired
   \begin{itemize}
-  \item Capabilities tagged with locality (local or global)   \note{We call non-local capabilities global}
-  \item New \emph{write-local} permission.
-  \item Local capabilities can only be stored by capabilities with \emph{write-local} permission
+  \item Capabilities tagged with locality (local or global)   %\note{We call non-local capabilities global}
+  \item New \textit{write-local} permission
+  \item Local capabilities can only be stored by capabilities with \textit{write-local} permission
   \end{itemize}
-  Calling convention highlights
-  \begin{itemize}
+  \note[item]<1->{To revoke a capability, we need to find it in memory which means we need access + need to search the entire memory.}
+  \note[item]<1->{Restricted where local capabilities can be stored. restricts where we need to look for a capability.}
+  \note[item]<1->{We define a calling convention. In order to prevent attack 1, we do the follwoing.}
+  \onslide<2->{Calling convention}
+  \begin{itemize}[<2->]
   \item Stack capability is local with permission read, write-local, and execute.
-  \item Clear stack before passing stack capability to untrusted code.
+    \note[item]<2->{Local stack capability cannot be stored on the heap. We need to be able to store old stack pointers somewhere, traditionally stack. }
+  \item No global write-local capabilities.
+    \note[item]<2->{Global write-local capabilities would undermine the entire idea as it would allow local capabilities to be stored indirectly.}
   \end{itemize}
@@
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \note{In the attack from before, when the attacker \emph{attempts to store the stack capability} on the heap, then the \emph{machine checks} that we have the correct authority to perform the operation.
+\emph{Assuming we only have global capabilities} for the heap, it cannot have write-local authority, due to the assumption on the previous slide, so we try to store the stack capability through a capability that does not have write-local authority, so it fails.}
+
     % recurrent parts
@@
       (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-    \draw node[right=0cm of sp1] { stack pointer };
+    \draw (11,14) node (a) { stack pointer };
+    \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
 
@@
     \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
-    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
+    \draw (14,8.25) edge[->,red,thick,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
     \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
@@
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \note[item]{While this prevents attack 1, we are not quite safe done.}
     % recurrent parts
     \stdstackstart
-
+    \note[item]{Trusted caller calls untrusted code.}
+    \note[item]<2->{untrusted code stores the stack pointer on the stack.}
+    \note[item]<2->{stack pointer local, but stack pointer has write local permission, so no problem.}
+    \note[item]<3->{untrusted code calls some trusted code with a callback.}
+    \note[item]<5->{trusted code runs for a bit pushes some local data to the stack and calls the callback.}
+    \note[item]<6->{The stack pointer is still on the stack allowing the untrusted code to read write to the stack frame of the trusted code.}
     % first step: capabilities
@@
         (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,12.5) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
       % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
         (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,12.5) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);        
     \end{onlyenv}
     \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
-    \draw<.> (6,13.25) edge[red,thick,->,in=45] (sp1);
+    \draw<.> (6,13.25) edge[red,thick,->,in=45] node[above right,yshift=0.1cm,xshift=-0.2cm] {\footnotesize local,\textsc{rwlx}}(sp1);
     \begin{onlyenv}<+->
@@
         (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (6,13.25) edge[red,thick,->,in=45] (spold);
+      \draw (6,13.25) edge[red,thick,->,in=45] node[sloped, above] {\footnotesize local,\textsc{rwlx}}(spold);
     \end{onlyenv}
@@
         (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+        \draw (11,12.5) node (a) { stack pointer };
+        \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
     \end{onlyenv}
@@
       (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+
+      \draw (11,12.5) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
     \end{onlyenv}
@@
   \end{tikzpicture}
-  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
+%  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
 \end{frame}
@@
   \frametitle{Calling Convention (Continued)}
+  \note[item]{Stack is basically the only place we can store local capabiliites.}
+  \note[item]{Make sure that untrusted code don't ``sneak'' capabilities between calls on the stack}
+  \note[item]{Clear stack and argument registers}
+  $\dots$
   \begin{itemize}
-  \item $\dots$
   \item Clear stack and non-argument registers before invoking untrusted code.
@@
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \note[item]{Let's see that the addition to the CC prevents attack 2.}
+    \note[item]{The untrusted code has been called. It calls the well-behaved code.}
+    \note[item]{The well-behaved code does its thing, but this time it clears the stack overwritting the old stack pointer the untrusted code had saved for later.}
+    \note[item]{The untrusted code starts running, but it does not have an old stack pointer available only the one given to them by the well-behaved code}
     % recurrent parts
@@
         (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,12.5) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);
     \end{onlyenv}
@@
       (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (11,12.5) node (a) { stack pointer };
+      \draw (a) edge[->,bend left] node[sloped,below] {\footnotesize local,\textsc{rwlx}} (sp1);      
     \end{onlyenv}
 
-    % attacker returns
-    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
-    \begin{onlyenv}<.>
-      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
+    % % attacker returns
+    % \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
+    % \begin{onlyenv}<.>
+    %   \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    %   \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+    %     (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+    %   \draw node[right=0cm of sp1] { stack pointer };
+    % \end{onlyenv}
 
-    % trusted code clears stack again
-    \begin{onlyenv}<+->
-      \foreach \x in {7,7.5,...,14}
-      {
-        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-    \end{onlyenv}
-    % then, the trusted code can return
-    \begin{onlyenv}<+>
-      \actadv{(0,4)}{(6,7)}{callee stack frame}
-    \end{onlyenv}
+    % % trusted code clears stack again
+    % \begin{onlyenv}<+->
+    %   \foreach \x in {7,7.5,...,14}
+    %   {
+    %     \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+    %   };
+    % \end{onlyenv}
+    % % then, the trusted code can return
+    % \begin{onlyenv}<+>
+    %   \actadv{(0,4)}{(6,7)}{callee stack frame}
+    % \end{onlyenv}
 
@@
   % and stack management
-  \note{We need to do a few more things, but this is the rest (just to illustrate that it is fairly simple).}
-  \note{Every point is motivated by some attack.}
+  \note[item]{The calling convention contains a bit more, but all of it is motivated by some attack.}
+  \note[item]{I won't motivate the rest here, but I wanted to show you that it does not take many more precautions.}
   \begin{itemize}
@@
     \begin{itemize}
-    \item Stack capabilitiy local capability with read, write-local, and execute authority.
-    \item No global write-local capabilities on the machine.
+    \item \textit{Stack capability local capability with read, write-local, and execute authority.}
+    \item \textit{No global write-local capabilities on the machine.}
     \end{itemize}
@@
     \begin{itemize}
-    \item Clear the stack.
+    \item \textit{Clear the stack.}
     \item Clear non-return registers.
@@
   \frametitle{Formalizing the Guarantees of a Capability Machine}
-  \note{How can we be sure calling convention works.
-Specifically, if a program interacts with intrusted code using the CC, can we formally show the correctness of the program if it relies on well-bracketedness or local-state encapsulation.}
-
-  \note{Need formal statement of the guarantees provided by the capability machine including the specific guarantees for local capabilities.}
+  \note[item]{How can we be sure the calling convention works?}
+  \note[item]{Specifically, if we have a \emph{program that interacts with untrusted code using the calling convention}, how do we formally show correctness of the program.}
+  \note[item]{We need a formal statement about the guarantees provided by the capabilities including the specific guarantees for local capabilities.}
+  \note[item]{Traditionally syntactic very syntactic (e.g. reference graph), does not take into account what the program does with its capabilities.}
+  \note[item]{\emph{We} have defined a logical relation which also give us a statement about the guarantees provided by the capability machine.}
   \begin{itemize}
-  \item How do we know the calling convention works?
-  \item Unary step-indexed Kripke logical relation over recursive worlds
+  \item How can we be sure the calling convention works?
+  \item<2-> Unary step-indexed Kripke logical relation over recursive worlds
     \begin{itemize}
-    \item Statement of guarantees probided by the capability machine
+    \item Statement of guarantees provided by the capability machine
     \end{itemize}
   \end{itemize}
-  \note{We state this formal statement in terms of a unary step-indexed Kripke logical relation over recursive worlds}
+  \note[item]<2->{Calling convention main application, but it is general}
 
-  \note{Calling convention main application, but it is very general - can be used to reason about other programs.}
+  \note[item]<2->{In the following: give some intuition about what a LR looks like for a capability machine}
 
-  \note{In the following: give some intuition for different parts of LR}
+  % \note{correctness here could be assert not violated}
 
-  \note{correctness here could be assert not violated}
-
-  \note{mention better than previous. Define the guarantees}
+  % \note{mention better than previous. Define the guarantees}
 \end{frame}
@@
     \begin{itemize}
-    \item<1-> Capabilities represent bound on executing code
+      \note[item]<1->{Compared to standard assembly language, capabilities executing code has access to represent bound.}
+    \item<1-> Capabilities represent bounds on executing code
+      \note[item]<1->{That is, the capabilities the executing code has access to.}
     \item<2-> World, $W$      
@@
 \end{columns}
-  % Do this as a series of figures?
+  % % Do this as a series of figures?
 
-  % Figure with memory and one capability to make the bound explicit
-  \note{compared to normal assembly, capabilities represent bounds on executing block of code.}
+  % % Figure with memory and one capability to make the bound explicit
+  % \note{compared to normal assembly, capabilities represent bounds on executing block of code.}
 
-  \note{we have no observable I/O, so the authority bounds we consider are related to memory. However, more fine-grained/detailed than read/write-authority.}
+  % \note{we have no observable I/O, so the authority bounds we consider are related to memory. However, more fine-grained/detailed than read/write-authority.}
 
-  % Color the memory to illustrate different regions of the world
-  \note{a piece of code can be bound by arbitrary memory invariants which we define in a \emph{world}.}
+  % % Color the memory to illustrate different regions of the world
+  % \note{a piece of code can be bound by arbitrary memory invariants which we define in a \emph{world}.}
 
-  % Write down a static invariant (some address contains 42, maybe write a legend on the left side of slide.).
-  \note{essentially, a world is a collection of invariants and safety of words defined with respect to a world.}
+  % % Write down a static invariant (some address contains 42, maybe write a legend on the left side of slide.).
+  % \note{essentially, a world is a collection of invariants and safety of words defined with respect to a world.}
 
-  \note{Define a set of words that are safe w.r.t. world W V(W) in P(Word)}
+  % \note{Define a set of words that are safe w.r.t. world W V(W) in P(Word)}
 
-  \note{Whether a capability is safe depends on the authority it carries}
+  % \note{Whether a capability is safe depends on the authority it carries}
 
-  % Add a read capability to the address that contains 42, safe
-  % Add read/write capability to the address that contains 42, not safe - can be used to violate the invariant.
-  \note{Example, w.r.t. a world with the invariant that an address contains constant. Safe for read capability, not write as write capability can break this invariant.}
+  % % Add a read capability to the address that contains 42, safe
+  % % Add read/write capability to the address that contains 42, not safe - can be used to violate the invariant.
+  % \note{Example, w.r.t. a world with the invariant that an address contains constant. Safe for read capability, not write as write capability can break this invariant.}
 
-  % Add a read capability that refers to itself
-  \note{safety for a read capability is the case if the read capability only gives access to safe capabilities.}
+  % % Add a read capability that refers to itself
+  % \note{safety for a read capability is the case if the read capability only gives access to safe capabilities.}
 
-  \note{What if that part of memory contains a read capability for the same part of memory? Cyclic definition. Solved by step-indexing.}
+  % \note{What if that part of memory contains a read capability for the same part of memory? Cyclic definition. Solved by step-indexing.}
 
-  \note{related to similar issue with languages with recursive types or higher-order ML-list references.}
+  % \note{related to similar issue with languages with recursive types or higher-order ML-list references.}
 
-  \note{Solved by step-indexing - safety up to a certain number of interaction steps.}
+  % \note{Solved by step-indexing - safety up to a certain number of interaction steps.}
 \end{frame}
@@
       f();
-      assert(x == 1))
+      assert(!x == 1))
     \end{lstlisting}
   \end{tabular}
-\note{example known from the litterature}
+\note{example known from the literature}
 \note{even just in ML difficult as f can be the closure.}
 \note{the assert can fail if the calls are not well-bracketed!}  
-\note{the local state is difficult to handle as the closure and the contex needs to be able to update the invariant for x in different ways.
+\note{the local state is difficult to handle as the closure and the context needs to be able to update the invariant for x in different ways.
 (closure can switch between 0 and 1 as it pleases, but context can transition only from 0 to 1.)}
 \note{relies heavily on well-bracketedness}
-\note{we have made a faithfull translation and proved correctness (i.e., the assertion never fails).}
+\note{we have made a faithful translation and proved correctness (i.e., the assertion never fails).}
 \note{more semantic statement of guarantees allows us to do this.}
@@
   \begin{itemize}
-  \item Capability machines can guarantee properties of high-level languages
-  \item Calling convention for well-bracketedness and local-state encapsulation
-  \item Unary step-indexed Kripke logical relation over recursive worlds
+  \item Capability machines can guarantee properties of high-level languages.
+  \item We developed a calling convention that ensures well-bracketedness and local-state encapsulation.
+  \item We define a unary step-indexed Kripke logical relation over recursive worlds.
     \begin{itemize}
-    \item Formal statement about guarantees provided by capability machine
-    \item Reasoning about programs in general
+    \item Formal statement about guarantees provided by capability machine.
+    \item Allows reasoning about programs on capability machine.
     \end{itemize}
-  \item Applied on the ``awkward example''
+  \item We apply it on the ``awkward example''.
   \end{itemize}

### 2018-04-15 presentations/esop18-04-2018/main.tex
more changes notes added to presentation

@@
     \begin{itemize}
-      \note[item]<1->{Compared to standard assembly language, capabilities executing code has access to represent bound.}
+      \note[item]<1-3>{Compared to standard assembly language, capabilities executing code has access to represent bound.}
     \item<1-> Capabilities represent bounds on executing code
-      \note[item]<1->{That is, the capabilities the executing code has access to.}
+      \note[item]<1-3>{That is, the capabilities the executing code has access to.}
+      \note[item]<1-3>{What exactly is the bound on. In the system we have considered, no I/O, so memory invariants.}
+      \note[item]<1-3>{Take what the program does into account - allows more fine-grained then simply ``read/write''}
     \item<2-> World, $W$
+      \note[item]<2-3>{The collection of invariants is what we call the world (and it can bethought of as  a model of the memory) }
+      \note[item]<2-3>{Here colored region represents invariant. A simple invariant could be a specific address contains 42.}
       \begin{itemize}
@@
       \end{itemize}
+      \note[item]<3-7>{We are interested in all the words on the machine that preserve these invariants, so}
     \item<4-> Predicate for safe values w.r.t world, $\stdvr(W)$
-      \begin{itemize}[<8->]
-      \item Recursively defined
+      \note[item]<4-7>{We define a predicate with theses values. Whether a capability violates invariants obviously depend on the invariants, so the predicate has to be world-indexed.}
+      \note[item]<4-7>{Let's see some examples:}
+      \note[item]<5-7>{If the capability on the slide has \emph{read} authority, then it cannot violate the simple invariant.}
+      \note[item]<6-7>{If the capability on the slide also has \emph{write} authority, then it can violate the invariant by simply overwriting that address with a different number. \emph{Not safe}.}
+      \note[item]<7-7>{Generally speaking a capability that can read is safe when it only can read safe words. What happens when it is stored at an address that it has authority over itself?}
+      \note[item]<7->{It is safe only if it can read only safe values which requires it to be safe.}
+      \begin{itemize}[<9->]
+      \item Recursively definition
+        \note[item]<8->{Need to take a fixed-point. Made possible by step-indexing.}
+        \note[item]<9->{Related to similar issue for languages with \emph{recursive types} and \emph{ML-like references}.}
       \end{itemize}
@@
     % One capability
-    \draw<1-6> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
+    \draw<1-7> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
     (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
@@
 
-    \fill<3-6>[black, fill opacity=0, text opacity=0.7] (0,4) rectangle (6,4.5) node[pos=.5] {\footnotesize 42};
+    \fill<3-7>[black, fill opacity=0, text opacity=0.7] (0,4) rectangle (6,4.5) node[pos=.5] {\footnotesize 42};
 
     \draw<5> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{r}} (sp1)   ;
-    \draw<6> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{rw}} (sp1)   ;
-    \draw<6>[very thick,color=red] (6.5,6.5) -- (8.5,4.5)
+    \draw<6-7> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{rw}} (sp1)   ;
+    \draw<7>[very thick,color=red] (6.5,6.5) -- (8.5,4.5)
     (6.5,4.5) -- (8.5,6.5);
 
-    \draw<7-> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
+    \draw<8-> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
     (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
-    \path<7->  (6,4.25) edge[black,out=30,in=135] node[above right]  {\footnotesize \textsc{r}} (8,5) 
+    \path<8->  (6,4.25) edge[black,out=30,in=135] node[above right]  {\footnotesize \textsc{r}} (8,5) 
               (8,5) edge[black,->,out=-45,in=45](sp1);
@@
   \begin{itemize}
+\note[item]<1->{Like languages with ML-references, memory changes over time. Example, if we are in this memory with two invariants and more is memory allocation. World need to cope with this.}
   \item<1-> Memory evolves over time
+    \note[item]<2->{Notion called \emph{future worlds}. Allow us to \emph{add invariants}.}
   \item<3-> Add invariants in future worlds
+    \note[item]<3->{Memory may be repurposed. Satic invariants don't model this. Alle invariants have an internal state machine. \emph{State machine progress in future worlds}.}
   \item<4-> Invariants as state machines
+    \note[item]<4->{Each state is associated with a predicate of all memories that respect the invariant.}
   \end{itemize}
@@
   \item<5-> Each state contains a predicate of accepted memory segments
+    \note[item]<5->{ We want to express ``all memories with safe values''. World dependent, so the predicate needs to be world indexed. }
   \item<6-> World indexed
+    \note[item]<6->{ Worlds with invariants with state machines with predicates that are world indexed - circular definition. }
+    \note[item]<6->{ Resolved using standard techniques from the litterature (essentially advanced step-indexing). }
   \end{itemize}
@@
 \end{overprint}
-  \note{Memory changes over time, for instance new memory may be allocated.}
-  \note{Allow worlds to evolve. New invariants can be added to handle freshly allocated memory.}
-  \note{Safety of words monotone w.r.t. worlds (which makes it into a Kripke Logical relation).}
+  % \note{Memory changes over time, for instance new memory may be allocated.}
+  % \note{Allow worlds to evolve. New invariants can be added to handle freshly allocated memory.}
+  % \note{Safety of words monotone w.r.t. worlds (which makes it into a Kripke Logical relation).}
 \end{frame}
@@
   \frametitle{Local Capabilities}
+  \note[item]<1>{Now consider how local capabilities affect all this.}
+  \note[item]<1>{Consider this simple example, first f is called with capability c as an argument. Then f is called with unit.}
+%  \note[item]<1>{What do we need to argue about c in the first call for it to be valid?}
+  \note[item]<1>{What may we assume about c in the second invocation of f?}
+  \note[item]<1>{Depends on c!}
+  \note[item]<2-3>{(Cont) Depends on c!}
   \texttt{f} is unknown code and \texttt{c} is a capability.
@@
   \item<2-> \texttt{c} global $\Rightarrow$ available in second invocation of \texttt{f}
-
-    \note{when c global, second invocation in a world where c is safe}
+    \note[item]<2-3>{If \emph{c is global}, then it can be \emph{stored on the heap}, so it needs to remain safe for the remainder of the execution. When f is invoked \emph{c must still be safe}.}
+%    \note{when c global, second invocation in a world where c is safe}
   \item<3-> \texttt{c} local $\Rightarrow$ not available in second invocation of \texttt{f}
-    \note{when c local, second invocation in a world where c is not (necessarily) safe}
-    \note{c essentially revoked, so the invariants it relies on need not hold.}
+    \note[item]<3>{If \emph{c local}, then \emph{CC dictates clear} all the places c may reside, so in the second invocation \emph{c need not remain safe}.}
+%    \note{when c local, second invocation in a world where c is not (necessarily) safe}
+%    \note{c essentially revoked, so the invariants it relies on need not hold.}
   \end{itemize}
-
-  \note{Two future world relations, one for all capabilities and one for non-local capabilities}
-  \note{If c global, then second invocation must happen in public future world, so c valid.}
-  \note{If c local, then second invocation may happen in a private future world.}
-  \begin{onlyenv}<4->
-  \begin{lemma}[Double monotonicity of value relation]
+  \note[item]<3>{Need two future world relations. In both, global capabilities must remain safe. In one local capabilities need not. We have public and private future world relation.}
+  \note[item]<4->{This double monotonicity lemma expresses the assumptions we can make, namely in \emph{public future world all capabilities} remain valid and in \emph{private future worlds local capabilities need not remain valid}.}
+  \note[item]<4->{In the example this means that if \emph{c is local}, then it is okay to invoke f a second time in a \emph{private future world} as c need not be safe anymore. }
+  \note[item]<4>{On the other hand, if \emph{c is global}, then the invokation of f must be in a \emph{public future world}, so c remains safe.}
+%  \note{Two future world relations, one for all capabilities and one for non-local capabilities}
+%  \note{If c global, then second invocation must happen in public future world, so c valid.}
+%  \note{If c local, then second invocation may happen in a private future world.}
+\onslide<4->{
+  \begin{lemma}[Double monotonicity of safety predicate]
     \begin{itemize}
@@
     \end{itemize}
-  \end{lemma}
-\end{onlyenv}
-  \note{How does local/global capabilities affect all this.}
-  \note{If we hand a global capability to untrusted code, then it may be stored in memory, so we will only be able to reinvoke that code if we can guarantee that those values are still valid. }
-  \note{Formally, the worlds contain the invariants that the global capability depend on and reinvocation is only possible in future worlds where these invariants are respected.}
-  \note{Local capabilities on the other provides a means to revoke capabilities.
+  \end{lemma}}
+\note[item]<4>{Related to public private future worlds, state machines with pub/priv transitions}
+%  \note{How does local/global capabilities affect all this.}
+%  \note{If we hand a global capability to untrusted code, then it may be stored in memory, so we will only be able to reinvoke that code if we can guarantee that those values are still valid. }
+%  \note{Formally, the worlds contain the invariants that the global capability depend on and reinvocation is only possible in future worlds where these invariants are respected.}
+%  \note{Local capabilities on the other provides a means to revoke capabilities.
 If we invoke untrusted code and give them a local capability, then they have no way to store it aside from the register file (and the stack), so when they return we can be sure that the local capability resides nowhere but the register file and the stack.}
-  \note{If we clear the register file and stack for old local capabilities before we invoke the untrusted code again, then we have essentially revoked those local capabilities and we can safely invoke the untrusted code again with the assumption that it does not have access to the local capability from the first call.}
-  \note{Formally, we allow this call to happen in a \emph{private future} world in which all global capabilities remain safe, but local capabilities don't.}
-  \note{We also have a \emph{public future} world relation in which \emph{all} capabilities remain safe.}
-  \note{Concretely, worlds may contain temporary regions on which local capabilities may depend.
+%  \note{If we clear the register file and stack for old local capabilities before we invoke the untrusted code again, then we have essentially revoked those local capabilities and we can safely invoke the untrusted code again with the assumption that it does not have access to the local capability from the first call.}
+%  \note{Formally, we allow this call to happen in a \emph{private future} world in which all global capabilities remain safe, but local capabilities don't.}
+%  \note{We also have a \emph{public future} world relation in which \emph{all} capabilities remain safe.}
+%  \note{Concretely, worlds may contain temporary regions on which local capabilities may depend.
 These regions persist in public future worlds which allows the local capabilities to rely on them.
@@
 
-  \note{include comment about priv/pub future worlds + interplay with monotonicity?}
+%  \note{include comment about priv/pub future worlds + interplay with monotonicity?}
 \end{frame}
@@
   \end{theorem}
+  \note[item]<1->{Now for the formal statement about guarantees}
+  \note[item]<1->{readCond: only safe values in the interval.}
+  \note[item]<1->{E-relation: when capability used as the pc with register file and memory respecting the world, then the execution respects the memory invariants.}
 
+  \note[item]<1->{In other words, take an arbitrary capability that only has access to safe values, then the memory invariants are preserved when we use it for execution.}
+  \note[item]<1->{The instructions it execute don't matter only the authority it can use.}
   % Comments/present high-level perspective
-  \note{readCond is the assumption that every thing in the interval [a,e] is safe to read.}
-  \note{$\stder$ is safe to execute relation. That is, it will respect all the memory invariants.}
-  \note{That is take an arbitrary capability. If it only has access to safe capabilities then it will preserve the invariants of the world.}
-  \note{Remeber, dynamic checks = failing is concidered secure} 
+  % \note{readCond is the assumption that every thing in the interval [a,e] is safe to read.}
+  % \note{$\stder$ is safe to execute relation. That is, it will respect all the memory invariants.}
+  % \note{That is take an arbitrary capability. If it only has access to safe capabilities then it will preserve the invariants of the world.}
+  % \note{Remeber, dynamic checks = failing is concidered secure} 
 \end{frame}
@@
   \end{tabular}
-\note{example known from the literature}
-\note{even just in ML difficult as f can be the closure.}
-\note{the assert can fail if the calls are not well-bracketed!}  
-\note{the local state is difficult to handle as the closure and the context needs to be able to update the invariant for x in different ways.
+  \note[item]{known from the litterature}
+  \note[item]{in ML difficult to reason about as callback f can be the closure it self. (so x can be either )}
+  \note[item]{assert may fail if calls not well-bracketed}
+  \note[item]{can do more things to attack well-bracketedness low-level. Context need not follow CC, so well-behaved code cannot rely on behavior of untrusted code. }
+  \note[item]{ We have proven a faithful translation correct.
+That is the assert never fails.
+Notice, dynamic checks, so machine can fail, but we set it up, so we can distinguish this from assertion failure.}
+  \note[item]{ More semantic statement of guarantees allow us to prove it. }
+% \note{example known from the literature}
+% \note{even just in ML difficult as f can be the closure.}
+% \note{the assert can fail if the calls are not well-bracketed!}  
+% \note{the local state is difficult to handle as the closure and the context needs to be able to update the invariant for x in different ways.
 (closure can switch between 0 and 1 as it pleases, but context can transition only from 0 to 1.)}
-\note{relies heavily on well-bracketedness}
-\note{we have made a faithful translation and proved correctness (i.e., the assertion never fails).}
-\note{more semantic statement of guarantees allows us to do this.}
+% \note{relies heavily on well-bracketedness}
+% \note{we have made a faithful translation and proved correctness (i.e., the assertion never fails).}
+% \note{more semantic statement of guarantees allows us to do this.}
 \end{frame}

### 2018-04-20 toplas/paper.tex
added toplas directory and tex files with suggestions for what to add

--- 2018-02-06 esop-2018/paper.tex
@@
-%% For double-blind review submission
-\documentclass{llncs}
-%% For single-blind review submission
-%\documentclass[acmsmall,10pt,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission
-%\documentclass[acmsmall10pt,]{acmart}\settopmatter{}
-
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format should change 'acmsmall' to
-%% 'sigplan'.
+\documentclass[format=acmsmall, review=false, screen=true]{acmart}
 
@@
                         %% http://ctan.org/pkg/booktabs
+
+% Metadata Information
+\acmJournal{TWEB}
+\acmVolume{9}
+\acmNumber{4}
+\acmArticle{39}
+\acmYear{2010}
+\acmMonth{3}
+\copyrightyear{2009}
+%\acmArticleSeq{9}
+% Copyright
+%\setcopyright{acmcopyright}
+\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\acmDOI{0000001.0000001}
+
+% Paper history
+%\received{February 2007}
+%\received[revised]{March 2009}
+%\received[accepted]{June 2009}
+
 \usepackage{subcaption} %% For complex figures with subfigures/subcaptions
@@
 
-\newcommand{\sectionname}{Section}
 \renewcommand{\figurename}{Figure}
@@
 
+\definecolor{OliveGreen}{RGB}{60,128,49}
+\newenvironment{toplas}
+    {\color{OliveGreen}
+
+    }{}
+\newcommand{\itoplas}[1]
+    {{\color{OliveGreen} #1}}
+
 \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
@@
 
-\bibliographystyle{MathPhySci}
+\bibliographystyle{ACM-Reference-Format}
+\citestyle{acmauthoryear}
 
 % Authors and affiliation
-\author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
-\institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
-\and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}
+%% Author with single affiliation.
+\author{Lau~Skorstengaard}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Aarhus University}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{Aarhus}
+  \state{State1}
+  \postcode{8210}
+  \country{Denmark}
+}
+\email{lau@cs.au.dk}          %% \email is recommended
+
+\author{Dominique~Devriese}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{iMinds-DistriNet}              %% \department is recommended
+  \institution{KU Leuven}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{-}
+  \state{-}
+  \postcode{-}
+  \country{Belgium}
+}
+\email{dominique.devriese@cs.kuleuven.be}          %% \email is recommended
+
+\author{Lars~Birkedal}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+%  \position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Aarhus University}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{Aarhus}
+  \state{State1}
+  \postcode{8210}
+  \country{Denmark}
+}
+\email{birkedal@cs.au.dk}          %% \email is recommended
+% \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
+% \institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
+% \and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}
 
@@
 attacks (leading to an attack-defense arms race) and/or rely on techniques like
-machine code rewriting~\cite{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
-verification~\cite{morrisett_system_1999}, virtual machines with a
-native stack~\cite{lindholm_java_2014} or
-randomization~\cite{forrest_building_1997}. The latter techniques essentially
+machine code rewriting~\citet{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
+verification~\citet{morrisett_system_1999}, virtual machines with a
+native stack~\citep{lindholm_java_2014} or
+randomization~\citep{forrest_building_1997}. The latter techniques essentially
 emulate protection techniques on existing hardware, at the cost of performance,
@@
 remediate these limitations with a better security model at
-the hardware level. They are based on old ideas~\cite{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
+the hardware level. They are based on old ideas~\citet{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
 but have recently received renewed interest; in particular, the CHERI
@@
 A similar problem is the enforcement of control-flow correctness on capability
-machines. An interesting approach is taken in CheriBSD~\cite{Watson2015Cheri}:
+machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
 the standard contiguous C stack is split into a central, trusted stack, managed
@@
   about programs on the capability machine. It is an untyped logical relation,
-  inspired by previous work on object capabilities \cite{Devriese:2016ObjCap}.
+  inspired by previous work on object capabilities \citet{Devriese:2016ObjCap}.
   We prove an analogue of the standard fundamental theorem of logical relations
@@
   a capability machine (a form of capability
-  safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
+  safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
   guarantees offered for local capabilities. It is very general and not tied to
@@
   \emph{single} orthogonal closure (rather than the earlier used biorthogonal
-  closure) and a variant of Dreyer et al.~\cite{Dreyer:jfp12}'s public and private future
-  worlds~\cite{Dreyer:jfp12} to express the special nature of local
+  closure) and a variant of \citet{Dreyer:jfp12}'s public and private future
+  worlds~\citep{Dreyer:jfp12} to express the special nature of local
   capabilities. The logical relation and the fundamental theorem expressing
@@
 For reasons of space, some details and all proofs have been omitted; please
-refer to the technical appendix~\cite{technical_appendix} for
+refer to the technical appendix~\citep{technical_appendix} for
 those.\dominique{upload TA as supplementary material instead?}
@@
 ($\rwlx$) and enter ($\entry$), which we will explain below.
+\itoplas{Suggestion: add the locality hierarchy and a short description. Maybe introduce the $\sqsubseteq$ relation here.}
 
@@
 \end{comment}
-\figurename~\ref{fig:op-sem} shows an excerpt of the operational semantics for a
+\figurename~\ref{fig:op-sem} \itoplas{Suggestion: extend this figure with more (all?) of the operational semantics.
+In general, should we add all the missing definitions and make this document closer to being self-contained.}
+shows an excerpt of the operational semantics for a
 few representative instructions. Essentially, a configuration $\Phi$ either
@@
 capability's permissions and global/local tag in the register file, but only if
-the new permissions are weaker than the original. It also never turns local
+the new permissions are weaker than the original
+%\itoplas{, which is what the relation $\sqsubseteq$ encapsulates.}
+.
+\itoplas{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
+%\itoplas{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
+% Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
+It also never turns local
 capabilities into global ones. $\geta{}{}$ queries the current address of a
@@
 further in the next section, but we refer to the technical
-appendix~\cite{technical_appendix} for full details.
+appendix~\citep{technical_appendix} for full details.
+\itoplas{Suggestion: add the malloc specification (maybe later as it uses the LR) and possibly a description of the linking tables. (Reviewer A, esop suggest explaining the linking model)}
+\itoplas{Suggestion: We also have the flag table ussed for assertions, maybe we should add a description of it here.}
+
+\itoplas{Reviewer C, popl, would have liked a discussion of the relationship between this capability machines and old capability machines from the litterature (Cambridge CAP \& IBM System 39, see tex comment).}
+ % * I'm old, so I'd appreciate a brief discussion LCM's relationship
+ %  with the capability hardware I (and perhaps other older readers, or
+ %  those outside this precise subfield) are likely to have heard of ---
+ %  the Cambridge CAP & IBM System 39.  I think the capabilities of the
+ %  machine are not that far from CAP, but this machine has the local
+ %  capabilities support. Is that right?
 
 \section{Stack and Return Pointer Management Using Local Capabilities}
-\label{sec:stack-and-return-pointer} One of the contributions in this paper is a
+\label{sec:stack-and-return-pointer}
+\itoplas{Suggestion: say something about arguments in this paragraph instead of defering it to the discussion only?}
+\itoplas{Suggestion: Make it more clear that this CC can be used to protect multiple components (reviewer A, ESOP), see tex comment.}
+\itoplas{Suggestion: Add some of our presentation illustration figures to illustrate some of the issues? May not be feasible as they rely on ``animation''.}
+% ## Clarify attacker model earlier
+% The proposed calling convention seems able to protect multiple
+% components from each other, yet most of the paper discusses this
+% protection only in terms of one single statically known
+% "adversary". This seems to be selling short the work, and raising
+% questions of whether this calling convention is much weaker
+% security-wise than the CHERI one, which is clearly about protecting
+% multiple components. It would be good to explain this better upfront,
+% not just in the discussions section at the end.
+\itoplas{Suggestion: Include a description of (possibly definition for) well-bracketedness?
+(requested by reviewer C, popl).
+Reviewer D, popl, asks for the same but also a definition of local state encapsulation.
+In the POPL notes, I wrote that I talked with Lars about it and this is something that is not easily defined.
+Perhaps we should have a discussion of this.}
+One of the contributions in this paper is a
 demonstration that local capabilities on a capability machine support a calling
@@
   give us temporary access to a callback not allocated on the stack.}
+\itoplas{Maybe a word on arguments passed from the adversary to us that we pass on to a callback given by the adversary (mentioned by reviewer E, popl, see tex comment.)}
+%page 8: "we only accept callbacks from the adversary in the form of global capabilities, which we dynamically check before invoking them".
+This makes sense, but don't you also need to check arguments that are passed from the adversary to us and then passed by us into the callback?
 
@@
 \label{para:macros}
+\itoplas{Add examples of macros to this section to make it clear what we mean by macros.}
+\itoplas{Move this subsection forward so it is just before the macros are used, i.e. the example section.}
+\itoplas{Reviewer C, ESOP, asks for high-level specifications for the macros in the style of malloc.
+I am not sure we want to actually include this in the submission, so it is maybe not here we should put our attention.
+(see review in tex comment}
+% Finally, as already mentioned above, the approach relies on `call` and `scall` macro-instructions that are only outlined in the submission, and described by very complicated pseudocode and English prose in the companion technical appendix.
+I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
 We define a number of reusable macros
@@
 
-The technical appendix~\cite{technical_appendix} contains detailed descriptions
+The technical appendix~\citep{technical_appendix} contains detailed descriptions
 of all the macros.
+\itoplas{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
 
@@
 \label{sec:logical-relation}
-
-In this section, we formalize the guarantees provided by the
-capability machine, including the specific guarantees for local
-capabilities, by means of a step-indexed Kripke logical relation 
-with recursively defined worlds. 
-We use the logical relation in the following section to show
-local-state encapsulation and control-flow integrity properties for
-challenging example programs.
-
+\itoplas{Suggestions for things to add heres
+  \begin{itemize}
+  \item Double monotonicity lemma (lemma 77 and 79) with a pointer back to local capability section of popl intro to this section.
+  \item A section with the lemmas we use for reasoning about programs
+    \begin{itemize}
+    \item lemma 58, scall works. Include description of how it handles the common things in call and leaves the argumentation for callee code and continuation.
+    \item lemma 59, malloc works. If this is included, we should probably also include the malloc specification.
+    \item lemma 60, create closure works.
+    \item Include a description about how the FTLR is used to reason about unknown code.
+    \item Mention that this is a step on the way towards having a program logic, and then in the discussion discuss why a program logic for this language is not what we need.
+    \end{itemize}
+  \item Reviewer B, ESOP suggests inferring general security properties from the LR.
+We had a section in the discussion that explained why this was not really possible.
+I have reintroduced it in the discussion.
+  \end{itemize}
+}
+
+% In this section, we formalize the guarantees provided by the
+% capability machine, including the specific guarantees for local
+% capabilities, by means of a step-indexed Kripke logical relation 
+% with recursively defined worlds. 
+% We use the logical relation in the following section to show
+% local-state encapsulation and control-flow integrity properties for
+% challenging example programs.
+
+\begin{toplas}
+\todo[inline]{Look this introductoin and the following section through to check for overlap.
+Also consider making backwards references to this introduction from the places where we use advanced techniques, so it becomes clear that they are necessary.}
 % Our logical relation will prescribe which values are capability safe given a
@@
 % will use.
-
-% Now that we have defined our calling convention, how can we be sure that it
-% works? More concretely, suppose that we have a program that uses the convention
-% in its interaction with untrusted adversary code. Can we formally prove the
-% program's correctness if it relies on well-bracketed control flow and private
-% state encapsulation for the interaction with the adversary? Clearly, such a
-% proof should depend on a formal expression of the guarantees provided by the
-% capability machine, including the specific guarantees for local capabilities.
-
-% In this section, we construct such a formalization. We make use of some
-% well-studied and powerful (but non-trivial) machinery from the literature.
-% Specifically, we employ a unary step-indexed Kripke logical relation with
-% recursive worlds, and some additional special characteristics of our own.
-% Step-indexing, Kripke logical relations and recursive worlds are techniques that
-% may be familiar from lambda calculus settings, but it may not be clear to the
-% reader how they apply in this more low-level assembly language. Therefore, we do
-% not immediately dive into the details, but we first try to provide some informal
-% intuition about how all of this machinery comes into play in our setting, in the
-% next section.
-
-% Note: even though the calling convention is the main application in this paper,
-% the logical relation we construct is very general and should be regarded as an
-% independent contribution.
-
-% \subsection{Formalizing the guarantees of the capability machine}
-% \label{sec:formalizing-guarantees}
-% What differentiates a capability machine from a more standard assembly language
-% is that we can bound the authority of an executing block of code, based solely
-% on the capabilities it has access to. Specifically, it does not matter which
-% instructions are actually executed, i.e., the bound also applies to untrusted
-% adversary code that has not been inspected or modified in any way.
-
-% \paragraph{Worlds}
-% But what does a ``bound on the authority'' of an executing block of code mean?
-% In our setting, there are no externally observable side-effects and the only
-% primitive authority that code may hold is authority over memory. As such, the
-% authority bounds we consider are related to memory, but in a form that is more
-% fine-grained than standard read/write authority: a piece of code's authority can be bounded
-% by arbitrary memory invariants that it is required to respect. Specifically, we
-% will define worlds $W \in \Worlds$, which describe a set of memory invariants, and
-% our results will express authority bounds on code as \emph{safety with respect to such
-% a world,} i.e., the fact that the code respects the invariants registered in
-% the world.
-
-% \paragraph{Safe values}
-% So, let's say that we have a world $W$ expressing that the memory must contain
-% value $42$ at address $0$, may contain arbitrary values at
-% addresses $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$, and
-% an integer at address $100$ that may only increase over time\footnote{Indeed, we
-%   will allow a notion of \emph{evolvable} invariants, aka \emph{protocols}, that can
-%   express such a temporal property.}. Our main theorem will state that if the
-% current register file only contains safe words (numbers or capabilities which
-% preserve the invariants in $W$ under any interaction), then an execution will
-% necessarily also preserve the memory invariants (irrespective of the instructions being
-% executed).
-
-% To make this more precise, we need to define the set $\stdvr(W) \in
-% \powerset{\Words}$ of words that are safe w.r.t.\ $W$. Essentially, the set
-% should only include words that preserve $W$'s invariants under any interaction,
-% but should otherwise be as liberal as possible. Numbers are clearly always safe,
-% as they cannot be used to break invariants. Whether a capability is safe depends
-% on the authority that it carries. In the above-described world, a read capability
-% for address $0$ is safe, as it can only be used to read the value $42$, which is
-% itself safe. However, a write capability for address $0$ is not safe: it can be
-% used to overwrite the memory at that address with a value other than $42$,
-% breaking the invariant for that address.
-
-% \paragraph{Step-indexing}
-% More generally, we want to define that a read capability for memory range
-% $[b,e]$ is safe, if the world guarantees that the words at those addresses are
-% themselves safe. However, this definition is cyclic: what if the world
-% guarantees that the memory at address $a$ will contain a read capability for
-% address $a$? The definition then just says that a read capability for address
-% $a$ is safe if and only if the same read capability for address $a$ is safe. This form of
-% cyclic reasoning is related to similar challenges in languages with recursive
-% types or higher-order ML-style references, and a standard solution is
-% to use step-indexing \cite{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
-% certain number of interaction steps. All words will be considered
-% safe up to $0$ steps (since if there is no interaction, nothing unsafe
-% can happen), and, for example, a read capability will be safe up to $n$
-% steps if the world guarantees that the words at the corresponding addresses are
-% safe up to $n - 1$ steps. We can then prove that the above read capability for
-% address $a$ is safe up to any number of steps.
-
-% \paragraph{Future worlds}
-% So worlds are defined as a set of invariants on the memory, but what if we
-% allocate fresh memory through malloc? We may want to establish new invariants
-% for this freshly allocated memory and be sure that the adversary will also
-% respect those (if we don't provide them with capabilities through which the new
-% invariants can be broken). To accommodate this, we allow worlds to evolve, for
-% example by adding additional invariants for freshly allocated memory. Formally,
-% we define valid ways for a world $W$ to evolve into a new world $W'$ through a
-% future-world relation $W' \future W$ and we ensure that the set of safe words in
-% world $W$ must remain safe in any future world $W'$. Defining safety w.r.t.\ a
-% notion of evolvable worlds makes our logical relation into a \emph{Kripke}
-% logical relation \cite{pitts_operational_1998}.
-
-% \paragraph{Invariants and Recursive Worlds}
-% So worlds group a set of memory invariants, but how are they actually
-% defined formally? We represent each invariant by a region
-% $r \in \Regions$. We will see later that regions contain state
-% machines to support a notion of evolvable invariants, but in every
-% state, they also contain a predicate $H$ that defines which memory
-% segments are acceptable in the current state of the invariant.
-% Unfortunately, it is not enough to just take
-% $H \in \powerset{\MemSegments}$, because sometimes the invariant may
-% itself be world-dependent. For example, we may want to express
-% invariants like ``the memory at address $50$ contains a value that is
-% safe in the current world''. As explained, worlds may evolve, and the
-% set of safe values may grow in future worlds, and therefore we need to
-% index $H$ over worlds, i.e., take
-% $H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
-% with worlds containing regions with world-indexed predicates,
-% i.e., the set of worlds must be recursively defined. We will see how such a recursive
-% definition can be accommodated using techniques from the
-% literature (essentially an advanced application of step-indexing).
-
-% \paragraph{Local capabilities}
-% A final thing we provide informal intuition for is how our results take into account local
-% capabilities and their special treatment by the hardware. Normally, when we
-% invoke an untrusted piece of code and provide it with certain global
-% capabilities, it may have stored those capabilities in memory and we will only
-% be able to reinvoke the code if we can guarantee that those values are still
-% valid. Formally, worlds represent the invariants that global capabilities'
-% safety relies on and the reinvocation is only safe in future worlds, where the
-% invariants are respected.
-
-% However, if we provide the adversary with local capabilities in that first
-% invocation, then the situation is a bit different. The adversary has no way to
-% store these local capabilities, so if we make sure that there are also no old
-% local capabilities in the register file for the second invocation (including the
-% capability being invoked), then the adversary cannot use them any more.%
-% \footnote{We are ignoring write-local capabilities in this discussion. If the
-%   adversary does have access to write-local capabilities in the first and second
-%   invocation, then the memory they address must entirely be cleared before the
-%   second invocation in order for the reinvocation to remain safe.} Therefore, we
-% can allow the second invocation to happen in \emph{private} future worlds ($W'
-% \futurestr W$), in which safe global capabilities remain safe, but local ones
-% don't. This private future world relation is more liberal than the standard,
-% public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
-% safe). Concretely, worlds may contain \emph{temporary} regions, representing
-% invariants that only local capabilities may rely on for their safety, and which
-% may be revoked (disabled) in private future worlds.
-
-% Interestingly, this idea is a variant of a notion of public/private future
-% worlds that has been previously used in the literature (see
-% \sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
-% and there is an interesting interplay with the recursiveness of the worlds: for
-% a temporary region, the predicate $H \in \Worlds \rightarrow
-% \powerset{\MemSegments}$ (which defines the safe memory segments in the
-% current world) is only required to be monotone w.r.t.\ public
-% future worlds (i.e.\ safe memory segments remain safe in public future worlds),
-% but for permanent regions, it must be monotone w.r.t.\ private future worlds.
-% As a consequence, the memory for a permanent region may not contain local
-% capabilities (as their safety would be broken in private future worlds), which
-% in turn implies that only local capabilities may have write-local permission (a
-% general sanity requirement when using local capabilities\footnote{In fact, local
-%   capabilities become useless as soon as the adversary has access to a single
-%   global, write-local capability.})\footnote{For explanation purposes, this
-%   discussion ignores certain ways to allow for local capabilities in a permanent
-%   region, for example, by not requiring that they are valid or requiring that
-%   they are local versions of valid global capabilities.}.
-
+Now that we have defined our calling convention, how can we be sure that it
+works? More concretely, suppose that we have a program that uses the convention
+in its interaction with untrusted adversary code. Can we formally prove the
+program's correctness if it relies on well-bracketed control flow and private
+state encapsulation for the interaction with the adversary? Clearly, such a
+proof should depend on a formal expression of the guarantees provided by the
+capability machine, including the specific guarantees for local capabilities.
+
+In this section, we construct such a formalization. We make use of some
+well-studied and powerful (but non-trivial) machinery from the literature.
+Specifically, we employ a unary step-indexed Kripke logical relation with
+recursive worlds, and some additional special characteristics of our own.
+Step-indexing, Kripke logical relations and recursive worlds are techniques that
+may be familiar from lambda calculus settings, but it may not be clear to the
+reader how they apply in this more low-level assembly language. Therefore, we do
+not immediately dive into the details, but we first try to provide some informal
+intuition about how all of this machinery comes into play in our setting, in the
+next section.
+
+Note: even though the calling convention is the main application in this paper,
+the logical relation we construct is very general and should be regarded as an
+independent contribution.
+
+\subsection{Formalizing the guarantees of the capability machine}
+\label{sec:formalizing-guarantees}
+What differentiates a capability machine from a more standard assembly language
+is that we can bound the authority of an executing block of code, based solely
+on the capabilities it has access to. Specifically, it does not matter which
+instructions are actually executed, i.e., the bound also applies to untrusted
+adversary code that has not been inspected or modified in any way.
+
+\paragraph{Worlds}
+But what does a ``bound on the authority'' of an executing block of code mean?
+In our setting, there are no externally observable side-effects and the only
+primitive authority that code may hold is authority over memory. As such, the
+authority bounds we consider are related to memory, but in a form that is more
+fine-grained than standard read/write authority: a piece of code's authority can be bounded
+by arbitrary memory invariants that it is required to respect. Specifically, we
+will define worlds $W \in \Worlds$, which describe a set of memory invariants, and
+our results will express authority bounds on code as \emph{safety with respect to such
+a world,} i.e., the fact that the code respects the invariants registered in
+the world.
+
+\paragraph{Safe values}
+So, let's say that we have a world $W$ expressing that the memory must contain
+value $42$ at address $0$, may contain arbitrary values at
+addresses $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$, and
+an integer at address $100$ that may only increase over time\footnote{Indeed, we
+  will allow a notion of \emph{evolvable} invariants, aka \emph{protocols}, that can
+  express such a temporal property.}. Our main theorem will state that if the
+current register file only contains safe words (numbers or capabilities which
+preserve the invariants in $W$ under any interaction), then an execution will
+necessarily also preserve the memory invariants (irrespective of the instructions being
+executed).
+
+To make this more precise, we need to define the set $\stdvr(W) \in
+\powerset{\Words}$ of words that are safe w.r.t.\ $W$. Essentially, the set
+should only include words that preserve $W$'s invariants under any interaction,
+but should otherwise be as liberal as possible. Numbers are clearly always safe,
+as they cannot be used to break invariants. Whether a capability is safe depends
+on the authority that it carries. In the above-described world, a read capability
+for address $0$ is safe, as it can only be used to read the value $42$, which is
+itself safe. However, a write capability for address $0$ is not safe: it can be
+used to overwrite the memory at that address with a value other than $42$,
+breaking the invariant for that address.
+
+\paragraph{Step-indexing}
+More generally, we want to define that a read capability for memory range
+$[b,e]$ is safe, if the world guarantees that the words at those addresses are
+themselves safe. However, this definition is cyclic: what if the world
+guarantees that the memory at address $a$ will contain a read capability for
+address $a$? The definition then just says that a read capability for address
+$a$ is safe if and only if the same read capability for address $a$ is safe. This form of
+cyclic reasoning is related to similar challenges in languages with recursive
+types or higher-order ML-style references, and a standard solution is
+to use step-indexing \citep{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
+certain number of interaction steps. All words will be considered
+safe up to $0$ steps (since if there is no interaction, nothing unsafe
+can happen), and, for example, a read capability will be safe up to $n$
+steps if the world guarantees that the words at the corresponding addresses are
+safe up to $n - 1$ steps. We can then prove that the above read capability for
+address $a$ is safe up to any number of steps.
+
+\paragraph{Future worlds}
+So worlds are defined as a set of invariants on the memory, but what if we
+allocate fresh memory through malloc? We may want to establish new invariants
+for this freshly allocated memory and be sure that the adversary will also
+respect those (if we don't provide them with capabilities through which the new
+invariants can be broken). To accommodate this, we allow worlds to evolve, for
+example by adding additional invariants for freshly allocated memory. Formally,
+we define valid ways for a world $W$ to evolve into a new world $W'$ through a
+future-world relation $W' \future W$ and we ensure that the set of safe words in
+world $W$ must remain safe in any future world $W'$. Defining safety w.r.t.\ a
+notion of evolvable worlds makes our logical relation into a \emph{Kripke}
+logical relation \citep{pitts_operational_1998}.
+
+\paragraph{Invariants and Recursive Worlds}
+So worlds group a set of memory invariants, but how are they actually
+defined formally? We represent each invariant by a region
+$r \in \Regions$. We will see later that regions contain state
+machines to support a notion of evolvable invariants, but in every
+state, they also contain a predicate $H$ that defines which memory
+segments are acceptable in the current state of the invariant.
+Unfortunately, it is not enough to just take
+$H \in \powerset{\MemSegments}$, because sometimes the invariant may
+itself be world-dependent. For example, we may want to express
+invariants like ``the memory at address $50$ contains a value that is
+safe in the current world''. As explained, worlds may evolve, and the
+set of safe values may grow in future worlds, and therefore we need to
+index $H$ over worlds, i.e., take
+$H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
+with worlds containing regions with world-indexed predicates,
+i.e., the set of worlds must be recursively defined. We will see how such a recursive
+definition can be accommodated using techniques from the
+literature (essentially an advanced application of step-indexing).
+
+\paragraph{Local capabilities}
+A final thing we provide informal intuition for is how our results take into account local
+capabilities and their special treatment by the hardware. Normally, when we
+invoke an untrusted piece of code and provide it with certain global
+capabilities, it may have stored those capabilities in memory and we will only
+be able to reinvoke the code if we can guarantee that those values are still
+valid. Formally, worlds represent the invariants that global capabilities'
+safety relies on and the reinvocation is only safe in future worlds, where the
+invariants are respected.
+
+However, if we provide the adversary with local capabilities in that first
+invocation, then the situation is a bit different. The adversary has no way to
+store these local capabilities, so if we make sure that there are also no old
+local capabilities in the register file for the second invocation (including the
+capability being invoked), then the adversary cannot use them any more.%
+\footnote{We are ignoring write-local capabilities in this discussion. If the
+  adversary does have access to write-local capabilities in the first and second
+  invocation, then the memory they address must entirely be cleared before the
+  second invocation in order for the reinvocation to remain safe.} Therefore, we
+can allow the second invocation to happen in \emph{private} future worlds ($W'
+\futurestr W$), in which safe global capabilities remain safe, but local ones
+don't. This private future world relation is more liberal than the standard,
+public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
+safe). Concretely, worlds may contain \emph{temporary} regions, representing
+invariants that only local capabilities may rely on for their safety, and which
+may be revoked (disabled) in private future worlds.
+
+Interestingly, this idea is a variant of a notion of public/private future
+worlds that has been previously used in the literature (see
+\sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
+and there is an interesting interplay with the recursiveness of the worlds: for
+a temporary region, the predicate $H \in \Worlds \rightarrow
+\powerset{\MemSegments}$ (which defines the safe memory segments in the
+current world) is only required to be monotone w.r.t.\ public
+future worlds (i.e.\ safe memory segments remain safe in public future worlds),
+but for permanent regions, it must be monotone w.r.t.\ private future worlds.
+As a consequence, the memory for a permanent region may not contain local
+capabilities (as their safety would be broken in private future worlds), which
+in turn implies that only local capabilities may have write-local permission (a
+general sanity requirement when using local capabilities\footnote{In fact, local
+  capabilities become useless as soon as the adversary has access to a single
+  global, write-local capability.})\footnote{For explanation purposes, this
+  discussion ignores certain ways to allow for local capabilities in a permanent
+  region, for example, by not requiring that they are valid or requiring that
+  they are local versions of valid global capabilities.}.
+\end{toplas}
 \subsection{Worlds}
@@
 high-level
-languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
+languages~\citep{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
 imposed by permanent regions stay in place indefinitely. Any capability, local
@@
 requirements on the memory. It simply serves as a mask for a revoked temporary
-region. Masking a region like this goes back to earlier work of Ahmed
-\cite{Ahmed2004semantics} and was also used by Birkedal et al.~\cite{Thamsborg:2011:KLR:2034773.2034831}.
+region. Masking a region like this goes back to earlier work of \citet{Ahmed2004semantics} and was also used by \citet{Thamsborg:2011:KLR:2034773.2034831}.
 
@@
 in Kripke models and the following lemma uses the method of
-Birkedal and Bizjak~\cite{Birkedal:2011:SKM:1926385.1926401}; Birkedal et al.~\cite{Birkedal:tutorial-notes} for
+\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
 constructing them. The formulation of the lemma is technical, so we recommend
@@
     \begin{aligned}
-      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{P} \ldotp \\
-      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
+      & \forall n' < n, W' \future W, a \in [\start',\addrend'] \subseteq [\start,\addrend], \perm \in \var{P} \ldotp \\
+      & \;\;\;\; \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\\
     \end{aligned} \right\} \\
@@
 precise definitions can be found in the technical
-appendix~\cite{technical_appendix}.
+appendix~\citep{technical_appendix}.
 
@@
 attempted, such as loading from a capability without read permission. This is
-similar to Devriese et al.~\cite{Devriese:2016ObjCap}'s logical relation for an untyped
+similar to \citet{Devriese:2016ObjCap}'s logical relation for an untyped
 language, but unlike typical logical relations for typed languages, which
@@
 
-The \emph{value relation} $\stdvr$ defines when words are safe. We make the
+The \emph{value relation} $\stdvr$ defines when words are safe.
+\itoplas{Explain why we use $\UPred{}$ with reference to the introduction of this section.}
+We make the
 value relation as liberal as possible by considering what is the most we can
@@
 $\iota^\pwl_{\start,\addrend}$ can be found
-in~\cite{technical_appendix}; it makes use of the isomorphism from
+in~\citep{technical_appendix}; it makes use of the isomorphism from
 Theorem~\ref{thm:world-existence}.
@@
 can be updated to point anywhere in their range, so they must be safe as a
-program counter (in the $\stder$-relation) no matter the current address. In
-contrast, enter capabilities are opaque and can only be used to jump to the
-address they point to. They also change permission when jumped to, so we require
+program counter (in the $\stder$-relation) no matter the current address.
+\itoplas{The range of an executable capability can also be updated, so they must also be safe as program counter no matter what their range of authority is reduced to.}
+In contrast, enter capabilities are opaque and can only be used to jump to the
+address they point to.
+\begin{toplas}
+  As enter capabilities are the only capabilities that have to use the address they point to, it is natural that the $\entryCond{}$ is also the only permission-based condition that depends on the current address of the capability.
+\end{toplas}
+They also change permission when jumped to, so we require
 them to be safe as a program counter after the permission is changed to $\exec$.
@@
 \end{theorem}
-The permission based conditions of Theorem~\ref{thm:ftlr} make sure that the
+The permission-based conditions of Theorem~\ref{thm:ftlr} make sure that the
 capability only provides safe authority in which case the capability must be in
@@
 guarantees offered by the capability machine, an instance of a general property
-called capability safety~\cite{Devriese:2016ObjCap,Maffeis2010OC}. To
+called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}. To
 understand this, consider that the theorem says the capability
@@
 examples depend on these properties for correctness.
-See the technical appendix~\cite{technical_appendix} for a detailed
+See the technical appendix~\citep{technical_appendix} for a detailed
 proof (by induction over the step-index $n$) of the theorem.
@@
 
+\itoplas{Suggestion: Add a small section about reasoning about programs.
+Add the ``scall works'' and ``malloc works'' lemmas and describe how they are used.
+This section could also point out the things one have to argue about that are hidden in similar high-level proofs (e.g., the return pointer valid when we return corresponding to restoring memory invariants of caller.).}
+
 \section{Examples}
 \label{sec:examples}
+\itoplas{Suggestion: add a couple of short proof sketches.}
 In this section, we demonstrate how our formalization of capability safety
@@
 unintelligible without the macros, as each macro expands to multiple basic instructions. The interested reader can find all the
-technical details in the technical appendix~\cite{technical_appendix}.
+technical details in the technical appendix~\citep{technical_appendix}.
 
@@
 pointer\footnote{These assumptions are kept intentionally vague for brevity.
-  Full statements are in the technical appendix~\cite{technical_appendix}.}.
+  Full statements are in the technical appendix~\citep{technical_appendix}.}.
 Furthermore
@@
 is a faithful translation of a tricky example known from the literature (known
-as the awkward example)~\cite{pitts_operational_1998,Dreyer:jfp12}. It consists
+as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists
 of two parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
@@
 complicated than the previous ones because it involves a closure invoked by the
-adversary and an adversary callback invoked by us. As explained in
+adversary and an adversary callback invoked by us.
+\itoplas{
+  Suggestion: Maybe give some more intuition as to why it is difficult.
+}
+As explained in
 \sectionname~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
@@
 
+\itoplas{Suggestion: Add an actual proof sketch}
 In Lemma~\ref{lem:correctness-g1} the order of control has been
@@
 \texttt{\footnotesize{f4}} returns.
-By utilizing the reinstation mechanism of the future world relation as well as our knowledge of the future worlds used, we can construct a world in which the adversary's invariants are preserved.
-The details of this and the proofs of the other lemmas are found in the technical appendix\cite{technical_appendix}.
-
-\begin{comment}
+
+\begin{toplas}
 In \figurename~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
@@
 have stashed away.
+\end{toplas}
 
@@
 \end{figure}
-\end{comment}
+
 \begin{comment}
@@
 \label{sec:discussion}
+\itoplas{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
+\itoplas{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
+% p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
+\itoplas{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}
+\itoplas{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
+% * Also, the local capabilities seem related to Boyland's "Borrowing"
+%   e.g. that is now supported in Rust amongst other research languages
+%   (Pony etc) also mentioned in the ECOOP paper.
+% * the combination of E and L capabilities seems very interesting
+%   indeed.  I wonder if e.g. they could be used to support strong
+%   models of encapsulation e.g. like ownership types. (see eg LNCS
+%   7850)
 \noindent\textbf{Calling convention}\\
@@
 that shows this, because it is not clear what such a theorem should look like.
-Formulations in terms of a control-flow graph, like the one by
-Abadi et al.~\cite{abadi_control-flow_2005}, do not take into account temporal properties,
+Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties,
 like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
@@
 prove that compiling such code to our machine using our calling convention is
-fully abstract~\cite{abadi_protection_1998}.
+fully abstract~\citep{abadi_protection_1998}.
+\itoplas{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
 
@@
 clear whether and how such a primitive could be implemented efficiently in the CHERI context.
+\itoplas{Maybe refer to some of the papers reviewer A, esop mentions for stack clearing. See comment in tex file. }
+% http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf
+% https://www.usenix.org/legacy/event/osdi08/tech/full_papers/zeldovich/zeldovich_html/index.html
+% http://users.ece.utexas.edu/~tiwari/pubs/MICRO-08-rangecache.pdf
 
@@
 
+\itoplas{We could also include a brief discussion of concurrency and the fact that this probably wouldn't scale to it (but well-bracktedness is not the same in that case anyway.) }
+\itoplas{Reviewer C, esop, wants to know what happens when there are several stack blocks (see tex for comment).
+We could here emphasize that while we talk about a single stack there could be multiple local read/write-local/execute-capabilities.
+If we get the stack from an untrusted source, which is the case in the awkward example, then we will use anything that \emph{looks like} the stack as the stack.
+However, due to the operational semantics for local capabilities all other ``stack blocks'' would have to be stored on the stack or in the register file.
+As the calling convention dictates that we clear these, the only place the alternative stack blocks will not be cleared is on a part of the stack that we do not have access to.
+The return capability would basically have to encapsulate this part of memory (perhaps indirectly) which means that from the moment we have been called, we will make sure there is only one stack available until our program returns from the initial call to it.}
+%p.10 "to prevent this from happening, we need to make sure the stack capability carries RWLX authority, since the system wide assumption then tells us that the adversary cannot have global capabilities to our local stack".
+It feels like the "L" capabitility is being overloaded to mean not just "this memory block has write-local authority and can contain local capabilities", but also to mean "this memory block is my stack and no other block is".
+What happens when there are several stack blocks, e.g.
+one per thread?
 \emph{Single stack} A single stack is a good choice for the simple capability
@@
 \\\\
-% \noindent \textbf{Reasoning about capability machine programs}\\
-% \emph{Semantic, but not syntactic properties} The logical relation defined in
-% \sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
-% programs. A limitation w.r.t. previous work is that the logical relation is
-% tailored exclusively towards semantic properties, not syntactic ones.
-
-% Imagine, for example, that we invoke a block of adversary code in such a way
-% that it only ever receives capabilities within a specific range of memory. After
-% the code returns, we may try to prove that any capabilities passed back to us in
-% the registers are still confined to that range of memory. The property of
-% falling in a certain address range is syntactic in the sense that it talks about
-% the specific implementation of a higher-order value rather than its behavior,
-% like the invariants that are required/preserved when we use it.
-
-% Such syntactic properties are hard to prove in our system. For the example
-% cited, it would be easy to conclude that the returned values are in the value
-% relation (see Figure~\ref{fig:logrel}). This gives us a lot of semantic
-% information, like conditions under which they are safe to use and invariants
-% that will be preserved when we do, but it does not tell us much about the
-% address of the capability. As a very concrete example, capabilities with
-% permission $\noperm$ are always in the value relation, irrespective of their
-% address. Semantically, this makes perfect sense, since they are always safe
-% since they cannot be used for anything anyway. However, it also means we do not
-% get syntactic information about them.
-
-% For our purposes, this restriction is unproblematic, since we are only
-% interested in proving semantic properties (e.g., an assertion will never fail).
-% However, in other situations, we may be interested in proving more syntactic
-% properties like the ones that are often considered in object capability
-% literature: confinement, no authority amplification etc. Although such
-% properties are more restrictive and less easy to use for reasoning,
-% \cite{Devriese:2016ObjCap} have demonstrated how a logical relation like
-% ours can be adapted to also support them, by quantifying the logical relation
-% over a custom interpretation of effectful computations and the type of
-% references. We expect their solution can be readily adapted to our setting,
-% modulo some details (like the fact that we do not just have read-write
-% capabilities, but also others).
+\begin{toplas}
+\noindent \textbf{Reasoning about capability machine programs}\\
+\emph{Semantic, but not syntactic properties} The logical relation defined in
+\sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
+programs. A limitation w.r.t. previous work is that the logical relation is
+tailored exclusively towards semantic properties, not syntactic ones.
+
+Imagine, for example, that we invoke a block of adversary code in such a way
+that it only ever receives capabilities within a specific range of memory. After
+the code returns, we may try to prove that any capabilities passed back to us in
+the registers are still confined to that range of memory. The property of
+falling in a certain address range is syntactic in the sense that it talks about
+the specific implementation of a higher-order value rather than its behavior,
+like the invariants that are required/preserved when we use it.
+
+Such syntactic properties are hard to prove in our system. For the example
+cited, it would be easy to conclude that the returned values are in the value
+relation (see Figure~\ref{fig:logrel}). This gives us a lot of semantic
+information, like conditions under which they are safe to use and invariants
+that will be preserved when we do, but it does not tell us much about the
+address of the capability. As a very concrete example, capabilities with
+permission $\noperm$ are always in the value relation, irrespective of their
+address. Semantically, this makes perfect sense, since they are always safe
+since they cannot be used for anything anyway. However, it also means we do not
+get syntactic information about them.
+
+For our purposes, this restriction is unproblematic, since we are only
+interested in proving semantic properties (e.g., an assertion will never fail).
+However, in other situations, we may be interested in proving more syntactic
+properties like the ones that are often considered in object capability
+literature: confinement, no authority amplification etc. Although such
+properties are more restrictive and less easy to use for reasoning,
+\citet{Devriese:2016ObjCap} have demonstrated how a logical relation like
+ours can be adapted to also support them, by quantifying the logical relation
+over a custom interpretation of effectful computations and the type of
+references. We expect their solution can be readily adapted to our setting,
+modulo some details (like the fact that we do not just have read-write
+capabilities, but also others).
+\end{toplas}
 % % \emph{No program logic} TODO: discuss this? 
-% \\\\
+\\\\
 \noindent\textbf{Logical relation}\\
@@
 existing pattern called biorthogonality. Biorthogonality is a pattern for
-defining logical relations~\cite{krivine_classical_1994,pitts_operational_1998}
+defining logical relations~\citet{krivine_classical_1994,pitts_operational_1998}
 in terms of an observation relation of safe configurations (like we do). The
@@
 
-Interestingly, Hur and Dreyer~\cite{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
+Interestingly, \citet{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
 Kripke logical relation for an assembly language (for reasoning about correct
@@
 code that relies on this (particularly, the original awkward example),
-Dreyer et al.~\cite{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
+\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
 contexts using Kripke worlds with public and private future world relations.
@@
 local capabilities are quite similar to a feature proposed for the high-level
-language Scala: Osvald et al.~\cite{osvald_gentrification_2016}'s second-class or local
+language Scala: \citet{osvald_gentrification_2016}'s second-class or local
 values. They are a kind of values that can be provided to other code for
@@
 
-Capability machines originate with Dennis and Van Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
-Levy~\cite{Levy1984capability} and Watson et al.~\cite{Watson2015Cheri} for an overview
+Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to \citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
 of previous work. The capability machine formalized in
@@
 correctness of control flow. This includes particularly the line of work on
-\emph{control-flow integrity}~\cite{abadi_control-flow_2005}. Those use a quite
+\emph{control-flow integrity}~\citep{abadi_control-flow_2005}. Those use a quite
 different attacker model than us: they assume an attacker that is not able to
@@
 secure compilation
-result \cite{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
+result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
 differs from theirs in that we use a different form of low-level security
@@
 
+\lau{Reviewer C, esop, would like a better comparison with \citec{Devriese:2016ObjCap} (but also admits that they did not follow the below discussion).}
 Our logical relation is a unary, step-indexed Kripke logical relation with
 recursive worlds
-\cite{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
-closely related to the one used by Devriese et al.~\cite{Devriese:2016ObjCap} to formulate
+\citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
+closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
 capability safety in a high-level JavaScript-like lambda calculus. Our
@@
 from previous work, like Kripke worlds with regions containing state transition
-systems \cite{Ahmed:popl09}, public/private future worlds \cite{Dreyer:jfp12}
+systems \citep{Ahmed:popl09}, public/private future worlds \citep{Dreyer:jfp12}
 (see \sectionname~\ref{sec:discussion} for a discussion), and biorthogonality
-\cite{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
-
-Swasey et al.~\cite{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
+\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
+
+\citet{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
 of object capability patterns. The logic is based on
-Iris~\cite{iris,iris2,iris3}, a state of the art higher-order concurrent
+Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
 separation logic and is formalized in Coq, building on the Iris Proof Mode for
-Coq~\cite{ipm}. OCPL gives a more abstract and modular way of proving
+Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
 capability safety for a lambda-calculus (with concurrency) compared to the
-earlier work by Devriese et al.~\cite{Devriese:2016ObjCap}.
-%domi: rest of the paragraph could be cut?
-% In the future we would also like to
-% investigate a new program logic for reasoning about capability safety for our
-% capability machine model. We think Iris would also be a natural starting point
-% for such an endeavour, since Iris is really a framework, which can be
-% instantiated to different programming languages. OCPL was able to leverage
-% existing Iris specifications for the high-level language; for our capability
-% machine model, however, it would be necessary to devise new kinds of
-% specifications for our low-level programs with unstructured control-flow. It is
-% likely that we could get inspiration from earlier work on logics for assembly
-% programming languages, such as XCAP~\cite{xcap}.
+earlier work by \citet{Devriese:2016ObjCap}.
+% domi: rest of the paragraph could be cut?
+\begin{toplas}
+In the future we would also like to
+investigate a new program logic for reasoning about capability safety for our
+capability machine model. We think Iris would also be a natural starting point
+for such an endeavour, since Iris is really a framework, which can be
+instantiated to different programming languages. OCPL was able to leverage
+existing Iris specifications for the high-level language; for our capability
+machine model, however, it would be necessary to devise new kinds of
+specifications for our low-level programs with unstructured control-flow. It is
+likely that we could get inspiration from earlier work on logics for assembly
+programming languages, such as XCAP~\citet{xcap}.
+\end{toplas}
+\itoplas{We could further add that this is not hte level of abstraction that one would like to reason.
+One would like to reason about the programs we actually write - that is programs written in high-level languages.
+If we had a fully-abstract compiler from a nice high-level language to this low-level machine, then we could reason about the program in the high-level language and then because the compiler is fully-abstract, we would retain all the nice properties from the high-level language.
+How does this work fit in to this picture?
+The calling convention ensures two properties that we expect our nice high-level languages to have, namely well-bracketedness and local-state encapsulation.
+If we want to do fully-abstract compilation, then we need some way to guarantee this after compilation.
+Further, logical relations are often used in full-abstractness proofs.
+This work expands our knowledge about logical relations should be defined for low-languages and in particular capability machines.
+(maybe also mention that enforcement necessary for fully-abstract compilation.
+Could be capability, but it could also be something else - we just need something which we do not have now.)  }
 
@@
 and uses it to prove a compartmentalization
-result~\cite{akram_el-korashy_formal_2016} (not implying control-flow
+result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
 correctness). He also adapts control-flow integrity (see above) to the machine
@@
 
-\subsubsection*{Acknowledgements.}
-This research was supported in part by the ModuRes Sapere Aude Advanced Grant
-from The Danish Council for Independent Research for the Natural Sciences (FNU).
-Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
+% \begin{acks}
+% This research was supported in part by the ModuRes Sapere Aude Advanced Grant
+% from The Danish Council for Independent Research for the Natural Sciences (FNU). 
+% Support for an STSM was received from COST Action EUTypes (CA15123).
+% Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
+% \end{acks}
 

### 2018-04-23 toplas/paper.tex
Fix some citet's that should be citep's.

@@
 attacks (leading to an attack-defense arms race) and/or rely on techniques like
-machine code rewriting~\citet{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
-verification~\citet{morrisett_system_1999}, virtual machines with a
+machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
+verification~\citep{morrisett_system_1999}, virtual machines with a
 native stack~\citep{lindholm_java_2014} or
@@
 remediate these limitations with a better security model at
-the hardware level. They are based on old ideas~\citet{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
+the hardware level. They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
 but have recently received renewed interest; in particular, the CHERI
@@
   about programs on the capability machine. It is an untyped logical relation,
-  inspired by previous work on object capabilities \citet{Devriese:2016ObjCap}.
+  inspired by previous work on object capabilities \citep{Devriese:2016ObjCap}.
   We prove an analogue of the standard fundamental theorem of logical relations
@@
 existing pattern called biorthogonality. Biorthogonality is a pattern for
-defining logical relations~\citet{krivine_classical_1994,pitts_operational_1998}
+defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998}
 in terms of an observation relation of safe configurations (like we do). The
@@
 likely that we could get inspiration from earlier work on logics for assembly
-programming languages, such as XCAP~\citet{xcap}.
+programming languages, such as XCAP~\citep{xcap}.
 \end{toplas}

### 2018-04-24 toplas/paper.tex
Minor modifications in toplas draft.

@@
 may be familiar from lambda calculus settings, but it may not be clear to the
-reader how they apply in this more low-level assembly language. Therefore, we do
+reader how they apply in this more low-level assembly language. Therefore, in the next section, we do
 not immediately dive into the details, but we first try to provide some informal
-intuition about how all of this machinery comes into play in our setting, in the
-next section.
+intuition about how all of this machinery comes into play in our setting.
 
@@
 \end{toplas}
-\itoplas{We could further add that this is not hte level of abstraction that one would like to reason.
+\itoplas{We could further add that this is not the level of abstraction that one would like to reason.
 One would like to reason about the programs we actually write - that is programs written in high-level languages.

### 2018-04-25 fully-abstract-comp/ftlr-proof-cases.tex
Finish another supporting proof for the ftlr about stackreadcond and fix minor problems I noticed here and there while making the proof.

@@
   From $\npair{(b,e)}\in\readCond{l,W}$, we get an $S \subseteq \addressable{\lin,
-    \pwheap}$, an $R : S \rightarrow \powerset{\mathbb{N}}$ with $\bigcup_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwheap(r)
-  \nsubeq \stdreg{R(r)}{\pur}$ for all $r \in S$.
+    \pwheap}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwheap(r)
+  \nsubeq \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
 
@@
   \begin{itemize}
-  \item $\ms_S =\ms_{f,S} \uplus \ms_S'$
+  \item $\ms_S =\ms_{S,f} \uplus \ms_S'$
   \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus
@@
   \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
-  Because $\pwheap(r) \nsubeq \stdreg{R(r)}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
-  also $\pwheap[W_\var{heap}](r) \nsubeq \stdreg{R(r)}{\pur}$ this
+  Because $\pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
+  also $\pwheap[W_\var{heap}](r) \nsubeq \stdreg{R(r),\gc}{\pur}$. This
   means that also $\npair{R_\ms(r)} \in
@@
 
-  Additionally, if $\npair{(b,e)}\in\writeCond{l,\pwheap}$, then
+  Additionally, if $\npair{(b,e)}\in\writeCond{l,W}$, then
   we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
-  for all $r \in S'$, $\pwheap(r) \nsupeq \stdreg{R'(r)}{\pur}$ and
+  for all $r \in S'$, $\pwheap(r) \nsupeq \stdreg{R'(r),\gc}{\pur}$ and
   $\pwheap(r) \text{ is address-stratified}$.
@@
   Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwheap[W_\var{heap}](r) \nsupeq \stdreg{R(r)}{\pur}$ and
-  $\pwheap[W_\var{heap}](r) \text{ is address-stratified}$.
+  $\pwheap[W_\var{heap}](r') \nsupeq \stdreg{R'(r'),\gc}{\pur}$ and
+  $\pwheap[W_\var{heap}](r') \text{ is address-stratified}$.
 
-  It follows that $r = r'$ because $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r) = \dom(R_\ms(r').2) = \dom(R_\ms(r').1)$ and all the $R_\ms(r).1$ and $R_\ms(r).2$ are disjoint.
+  It follows that $r = r'$ because $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r) \ni a$ and $\dom(R_\ms(r').2) = \dom(R_\ms(r').1) = R'(r') \ni a$ and all the $R_\ms(r).1$ and $R_\ms(r).2$ are disjoint.
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupeq[n-1] \stdreg{R(r)}{\pur}$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupeq \stdreg{R'(r),\gc}{\pur}$.
 
   From this, it follows that $\memSatHeap{\ms_S\update{a}{0},\ms_\var{T,heap}\update{a}{0}}{W_r' \oplus W_\var{heap}'}$ and finally $\memSat{(\ms_S\update{a}{0},\stk,\ms_\stk,\ms_T\update{a}{0})}{W_M'}$.
+\end{proof}
+
+\begin{lemma}[stackReadCondition works]
+  \label{lem:readcond-writecond-work}
+  If
+  \begin{itemize}
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
+  \item $\npair{(b,e)}\in\stackReadCond{W}$
+  \item $a \in [b,e]$
+  \item $W \oplus W_M$ is defined
+  \item $n' < n$
+  \end{itemize}
+  Then $\npair[n']{(\ms_\stk(a),\ms_T(a))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
+  Additionally, if
+  \begin{itemize}
+  \item $\npair{(b,e)}\in\stackWriteCond{W}$
+  \end{itemize}
+  Then $\memSat{(\ms_S,\stk,\ms_\stk[a\mapsto 0],\ms_T[a\mapsto 0])}{W_M'}$.
+\end{lemma}
+\begin{proof}
+  From $\npair{(b,e)}\in\stackReadCond{W}$, we get an $S \subseteq \addressable{\lin,
+    \pwfree}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwfree(r)
+  \nsubeq \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
+
+  Since $a \in [b,e]$, there is a unique $r \in S$ such that $a \in R(r)$.
+
+  Since $W \oplus W_M$ is defined, we have that $r \in \dom(\pwfree) =
+  \dom(\pwfree[W_M])$ and $\pwfree(r) \oplus \pwfree[W_M](r)$ is defined.
+
+  From $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$, we get that
+  $\stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m)$,
+  $\ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m$ is defined,
+  $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
+  $\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}$, $\ms_{T,f}$, $\ms_{S,f}$, $\ms_S'$ such that
+  \begin{itemize}
+  \item $\ms_S =\ms_{S,f} \uplus \ms_S'$
+  \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus
+    \ms_\var{T,heap} \uplus \ms_{T,f}$
+  \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
+  \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
+  \item $\memSatHeap{\ms_S',\ms_\var{T,heap}}{W_\var{heap}}$.
+  \end{itemize}
+
+  From $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$, we get an
+  $R_\ms : \dom(\activeReg{\pwfree[W_\var{free\_stack}]}) \fun \MemSeg \times \MemSeg$,
+  $\ms_\var{T,free\_stack} = \biguplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} \pi_2(R_\ms(r))$,
+  $\ms_\stk = \biguplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} \pi_1(R_\ms(r))$,
+  $\stkb \in \dom(\ms_\var{T,free\_stack}) \wedge \stkb \in \dom(\ms_\stk)$,
+  $\exists R_W : \dom(\activeReg{\pwfree[W_\var{free\_stack}]}) \fun \World\ldotp$
+  $W_\var{free\_stack} = \oplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} R_W(r)$,
+  $\forall r \in \activeReg{\pwfree[W_\var{free\_stack}]}$, we have that
+  $\npair{R_\ms(r)} \in  \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
+
+  We have that $r \in \addressable{\lin, \pwfree} \subseteq
+  \activeReg{\pwfree[W_\var{free\_stack}]}$, so $\npair{R_\ms(r)} \in
+  \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
+  Because $\pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{free\_stack} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
+  also $\pwfree[W_\var{free\_stack}](r) \nsubeq \stdreg{R(r),\gc}{\pur}$. This
+  means that also $\npair{R_\ms(r)} \in
+  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined $\nsubeq$ somewhere?}
+
+  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r)$ and we get a $S : R(r) \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in R(r)} S(\aaddr)$ and $\forall \aaddr \in R(r),n' < n \ldotp \npair[n']{(\ms_\stk(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
+
+  Since $\aaddr \in R(r)$ and $n' < n$, we have that $\npair[n']{(\ms_\stk(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
+  we can take $W_M' = W_r' \oplus W_{\var{heap}} \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}}')$ with $W_r' = \oplus_{\aaddr \in (R(r)\setminus \{\aaddr\})} S(\aaddr)$ and $W_\var{free\_stack}' = \oplus_{r' \in (\dom(\activeReg{\pwheap[W_\var{heap}]})\setminus \{r\})} R_W(r')$, and get
+  \begin{align*}
+    S(\aaddr) \oplus W_M'
+    &=S(\aaddr) \oplus (W_r' \oplus W_{\var{free\_stack}}' \oplus (W_{\var{stack}}\oplus W_\var{heap}))\\
+    &=
+    \oplus_{\aaddr \in R(r)} S(\aaddr) \oplus W_{\var{free\_stack}}' \oplus (W_{\var{stack}}\oplus W_{\var{heap}})\\
+    &=
+    \xi(\xi^{-1}(R_W(r))) \oplus W_{\var{free\_stack}}' \oplus (W_{\var{stack}} \oplus W_\var{heap})\\
+    &=
+    R_W(r) \oplus W_{\var{free\_stack}}' \oplus (W_{\var{stack}} \oplus W_\var{heap})\\
+    &=
+    \oplus_{r \in \dom(\activeReg{\pwheap[W_\var{free\_stack}]})} R_W(r)\oplus (W_{\var{stack}} \oplus W_{\var{heap}})\\
+    &=
+      W_\var{free\_stack} \oplus (W_{\var{stack}} \oplus W_{\var{heap}})\\
+      &= W_M
+  \end{align*}
+
+  Additionally, if $\npair{(b,e)}\in\stackWriteCond{W}$, then
+  we get an $S'\subseteq \addressable{\lin, \pwfree} \subseteq \activeReg{\pwheap[W_\var{free\_stack}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ and
+  for all $r \in S'$, $\pwfree(r) \nsupeq \stdreg{R'(r),\gc}{\pur}$ and
+  $\pwfree(r) \text{ is address-stratified}$.
+
+  Since $a \in [b,e]$, there is an $r' \in S'$ such that $a \in R'(r')$ .
+
+  Because $W \oplus W_M$ is defined, it follows that also 
+  $\pwfree[W_\var{free\_stack}](r') \nsupeq \stdreg{R'(r'),\gc}{\pur}$ and
+  $\pwfree[W_\var{free\_stack}](r') \text{ is address-stratified}$.
+
+  It follows that $r = r'$ because $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r) \ni a$ and $ \dom(R_\ms(r').2) = \dom(R_\ms(r').1) = R'(r') \ni a$ and all the $R_\ms(r).1$ and $R_\ms(r).2$ are disjoint.
+
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r) \nsupeq \stdreg{R'(r),\gc}{\pur}$.
+
+  From this, it follows that $\memSatFStack{\ms_\stk\update{a}{0},\ms_\var{T,heap}\update{a}{0}}{W_r' \oplus W_\var{free\_stack}'}$ and finally $\memSat{(\ms_S,\stk,\ms_\stk\update{a}{0},\ms_T\update{a}{0})}{W_M'}$.
 \end{proof}
@@
 \begin{proof}
+  From $\npair{(c,c')}\in\lrv(H_\sigma,W)$ with $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$, $\perm \in \readAllowed{}$ and $\perm' \in \readAllowed{}$, we get that $\perm = \perm'$, $\lin' = \linear$, $\baddr = \baddr'$, $\eaddr = \eaddr'$ and $\aaddr = \aaddr'$ and $\npair{(\baddr,\eaddr)} \in \stackReadCond{W}$.
+
+  TODO: fix rest of proof
+
+  Lemma~\ref{lem:readcond-writecond-work} then gives us a $W'$ and $W_M'$ such that $W_M = W' \oplus W_M'$ and $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$.
+
+  It remains to prove that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+  We have to distinguish the case that $\isLinear{\ms_S(\aaddr)}$ and the opposite case.
+  \begin{itemize}
+  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
+    From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
+    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+
+  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
+    The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by non-expansiveness of memory satisfaction, i.e.
+Lemma~\ref{lem:non-expansive}.
+  \end{itemize}
   TODO
@@
 From the operational semantics and in particular the semantics of $\scall{}{}{}{}$, we know $\Phi_S(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\npair[n']{(\src{\reg_S(\rstk)},\trg{\reg_T(\rstk)})} \in \lrv(W_{R,\stk})$.
-From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree[W_{R,\stk}](r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk']}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
+From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree[W_{R,\stk}](r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk'],\gc}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
 Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}.
@@
     W' = W     & \update{\mathrm{free}.r}{\revoked}\\
-               & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1]}{\spa}}\\
-               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk]}{\spa},\opc)}\\
-               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\eaddr_\stk']}{\spa}}
+               & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1],\gc}{\spa}}\\
+               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk],\gc}{\spa},\opc)}\\
+               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\eaddr_\stk'],\gc}{\spa}}
   \end{array}

### 2018-04-25 fully-abstract-comp/technical-report.tex
added missing i in world specifies component.
Fix use of wrong macro in nsubeq def.
Small details.
Finish another supporting proof for the ftlr about stackreadcond and fix minor problems I noticed here and there while making the proof.

@@
 
+\dominique{Remaining question: what exactly does $\nsubeq$ on a UPred mean?  Does it mean subset ``up to n (including)'' or ``up to n-1 (including)''?
+  Check what we can require without breaking contractiveness of recursive definition of value relation and/or what we did in previous paper and/or what we need in FTLR proof.
+  Update: in the FTLR proof, we seem to need ``up to n'' because the heap predicate of the standard region is contractive in terms of the value relation.
+  This also means we can afford to use ``up to n'' without breaking contractiveness of the recursive definition of the value relation.
+  }
+We have the following notion of regions being an $n$-subset of another region.
+\begin{mathpar}
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
+            { (v,H) \nsubeq (v',H') }
+\and
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
+            { (\pure,H,H_\sigma) \nsubeq (v',H') }
+\and
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
+            { (v,H) \nsubeq (\pure,H',H_\sigma) }
+\and
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
+            { (\pure,H,H_\sigma) \nsubeq (\pure,H',H_\sigma') }
+\end{mathpar}
+We define $n$-equality of regions $\iota \nequal \iota'$ as the symmetric closure, i.e. $\iota \nequal \iota'$ iff $\iota \nsubeq \iota'$ and $\iota \nsupeq \iota'$.
+
 % \[
@@
       \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
-      \quad \exists M : S \fun \powerset{\MemSeg}, R : S \fun \powerset{\Seal}, C : S \fun \powerset{\Seal} \\
+      \quad \exists M : S \fun \powerset{\MemSeg} \\
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
-      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq[n-1] \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}
@@
 \[
-  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc)
+  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc) , v \in \{\spa,\spao\}
 \]
 for readability, we use $\spao$ short for $\spatialo$, $\spa$ short for $\spatial$, and $\pur$ as short for $\pure$.
+\[
+  \stdreg{A,\gc}{\pur} \defeq (\pur,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
+\]
 \lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
@@
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-    \dom(\code) \nsubseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
+    \dom(\code) \subseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
@@
   \begin{itemize}
-  \item $\pwheap(r_i) = \codereg{\sigrets_i,\sigcloss_i,\code_i,\gc}$
+  \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
   \item $\mscode = \biguplus_i \code_i$

### 2018-04-25 toplas/paper.tex
Added subseg and lea to table of instructions.

@@
 \newcommand{\perm}{\var{perm}}
+\newcommand{\inftyend}{-42}
 \newcommand{\permp}{\var{permPair}}
@@
 \newcommand{\decodePerm}[1]{\plainfun{decodePerm}{#1}}
+\newcommand{\decodeEnd}[1]{\plainfun{decodeEndAddr}{#1}}
+\newcommand{\decodeStart}[1]{\plainfun{decodeStartAddr}{#1}}
 \newcommand{\updatePcPerm}[1]{\plainfun{updPcPerm}{#1}}
@@
     \hline
+    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
+                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
+    \hline
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
+    \hline
     $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$ and $(\perm',g') = \decodePermPair{\memreg(r_2)}$
@@
     \hline
+    \itoplas{$\subseg{\refreg{r_1}}{r_2}{r_3}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{w}}}$} &  \itoplas{$\memreg(r_1) = \stdcap$ and for $i \in \{2,3\}$ $n_i = \memreg(\rv_i)$ and $n_2 \in \nats$ and $\start \leq n_2$ and $n_3 \leq \addrend$ where either $n_3 \in \nats$ or ($n_3=\inftyend$ and $\addrend = \infty$) and $\perm \neq \entry$ and $w = ((\perm,\gl),n_1,n_2,\addr)$}\\
+    \hline
+    \itoplas{$\lea{\refreg{r_1}}{r_2}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{c}}}$} &\itoplas{$\memreg(r_1) = \stdcap$ and $n = \memreg(r_2)$ and $n \in \ints $ and $\perm \neq \entry$ and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$} \\
+    \hline
     $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
                                                 $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
-    \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
-    \hline
-    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
-                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
     \hline

### 2018-04-25 toplas/paper.tex
Added new command for suggestions that has different colour than addition.

@@
 \definecolor{OliveGreen}{RGB}{60,128,49}
+\definecolor{Blue}{RGB}{92,26,143}
 \newenvironment{toplas}
@@
     }{}
+\newenvironment{toplassug}
+    {\color{Blue}
+
+    }{}
+
 \newcommand{\itoplas}[1]
     {{\color{OliveGreen} #1}}
+\newcommand{\itoplassug}[1]
+    {{\color{Blue} #1}}
 
@@
 \end{comment}
-\figurename~\ref{fig:op-sem} \itoplas{Suggestion: extend this figure with more (all?) of the operational semantics.
+\figurename~\ref{fig:op-sem}
+%\itoplassug{Suggestion: extend this figure with more (all?) of the operational semantics.
 In general, should we add all the missing definitions and make this document closer to being self-contained.}
@@
 read permission and an appropriate range). $\restricttwo{}{}$ updates a
-capability's permissions and global/local tag in the register file, but only if
-the new permissions are weaker than the original
-%\itoplas{, which is what the relation $\sqsubseteq$ encapsulates.}
-.
-\itoplas{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
-%\itoplas{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
+capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original. 
+%\itoplassug{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
+%\itoplassug{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
 % Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
@@
 appendix~\citep{technical_appendix} for full details.
-\itoplas{Suggestion: add the malloc specification (maybe later as it uses the LR) and possibly a description of the linking tables. (Reviewer A, esop suggest explaining the linking model)}
-\itoplas{Suggestion: We also have the flag table ussed for assertions, maybe we should add a description of it here.}
-
-\itoplas{Reviewer C, popl, would have liked a discussion of the relationship between this capability machines and old capability machines from the litterature (Cambridge CAP \& IBM System 39, see tex comment).}
+further in the next section.
+\itoplassug{Suggestion: We also have the flag table ussed
+  for assertions, maybe we should add a description of it here.\\}
+\itoplassug{Reviewer C, popl, would have liked a discussion of the relationship
+  between this capability machines and old capability machines from the
+  litterature (Cambridge CAP \& IBM System 39, see tex comment).\\}
  % * I'm old, so I'd appreciate a brief discussion LCM's relationship
@@
 \label{sec:stack-and-return-pointer}
-\itoplas{Suggestion: say something about arguments in this paragraph instead of defering it to the discussion only?}
-\itoplas{Suggestion: Make it more clear that this CC can be used to protect multiple components (reviewer A, ESOP), see tex comment.}
-\itoplas{Suggestion: Add some of our presentation illustration figures to illustrate some of the issues? May not be feasible as they rely on ``animation''.}
+\itoplassug{Suggestion: say something about arguments in this paragraph instead of defering it to the discussion only?\\}
+\itoplassug{Suggestion: Add some of our presentation illustration figures to illustrate some of the issues? May not be feasible as they rely on ``animation''.\\}
 % ## Clarify attacker model earlier
@@
 % not just in the discussions section at the end.
-\itoplas{Suggestion: Include a description of (possibly definition for) well-bracketedness?
+\itoplassug{Suggestion: Include a description of (possibly definition for) well-bracketedness?
 (requested by reviewer C, popl).
@@
 In the POPL notes, I wrote that I talked with Lars about it and this is something that is not easily defined.
-Perhaps we should have a discussion of this.}
+Perhaps we should have a discussion of this.\\}
 One of the contributions in this paper is a
@@
   give us temporary access to a callback not allocated on the stack.}
-\itoplas{Maybe a word on arguments passed from the adversary to us that we pass on to a callback given by the adversary (mentioned by reviewer E, popl, see tex comment.)}
+\itoplassug{Maybe a word on arguments passed from the adversary to us that we pass on to a callback given by the adversary (mentioned by reviewer E, popl, see tex comment.)}
 %page 8: "we only accept callbacks from the adversary in the form of global capabilities, which we dynamically check before invoking them".
@@
 of all the macros.
-\itoplas{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
+%\itoplassug{Suggestion: Make it more clear that this CC can be used to protect multiple components (reviewer A, ESOP), see tex comment.\\}
+\begin{toplas}
+  \textbf{Modularity} The calling convention ensures well-bracketed calls and local-state encapsulation for the caller, but not the callee.
+In the above presentation to make it easy to distinguish between the parties involved, we present the callee as some adversarial code that we do not trust.
+In reality, the callee could be well-behaved and wish to ensure well-bracketed calls and local-state encapsulation.
+The calling convention puts no restriction on the callee that the caller itself does not follow, so by following the calling convention, the callee would also get well-bracketedness and local-state encapsulation.
+Generally, the calling convention ensures the two properties for anyone that employ it, so it scales to the setting where multiple distrusting parties need to invoke each other and each of them can protect themself by using the calling convention.
+\end{toplas}
 
@@
 \label{sec:logical-relation}
-\itoplas{Suggestions for things to add heres
+\itoplassug{Suggestions for things to add heres
   \begin{itemize}
@@
 The \emph{value relation} $\stdvr$ defines when words are safe.
-\itoplas{Explain why we use $\UPred{}$ with reference to the introduction of this section.}
+\itoplassug{Explain why we use $\UPred{}$ with reference to the introduction of this section.}
 We make the
@@
 program counter (in the $\stder$-relation) no matter the current address.
-\itoplas{The range of an executable capability can also be updated, so they must also be safe as program counter no matter what their range of authority is reduced to.}
+\itoplassug{The range of an executable capability can also be updated, so they must also be safe as program counter no matter what their range of authority is reduced to.}
 In contrast, enter capabilities are opaque and can only be used to jump to the
@@
 
-\itoplas{Suggestion: Add a small section about reasoning about programs.
+\itoplassug{Suggestion: Add a small section about reasoning about programs.
 Add the ``scall works'' and ``malloc works'' lemmas and describe how they are used.
@@
 \label{sec:examples}
-\itoplas{Suggestion: add a couple of short proof sketches.}
+\itoplassug{Suggestion: add a couple of short proof sketches.}
 In this section, we demonstrate how our formalization of capability safety
@@
 adversary and an adversary callback invoked by us.
-\itoplas{
+\itoplassug{
   Suggestion: Maybe give some more intuition as to why it is difficult.
@@
 
-\itoplas{Suggestion: Add an actual proof sketch}
+\itoplassug{Suggestion: Add an actual proof sketch}
 In Lemma~\ref{lem:correctness-g1} the order of control has been
@@
 \label{sec:discussion}
-\itoplas{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
-\itoplas{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
+\itoplassug{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
+\itoplassug{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
 % p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
-\itoplas{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}
-\itoplas{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
+\itoplassug{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}
+\itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
 % * Also, the local capabilities seem related to Boyland's "Borrowing"
@@
 fully abstract~\citep{abadi_protection_1998}.
-\itoplas{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
+\itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
 
@@
 clear whether and how such a primitive could be implemented efficiently in the CHERI context.
-\itoplas{Maybe refer to some of the papers reviewer A, esop mentions for stack clearing. See comment in tex file. }
+\itoplassug{Maybe refer to some of the papers reviewer A, esop mentions for stack clearing. See comment in tex file. }
 % http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf
@@
 
-\itoplas{We could also include a brief discussion of concurrency and the fact that this probably wouldn't scale to it (but well-bracktedness is not the same in that case anyway.) }
-\itoplas{Reviewer C, esop, wants to know what happens when there are several stack blocks (see tex for comment).
+\itoplassug{We could also include a brief discussion of concurrency and the fact that this probably wouldn't scale to it (but well-bracktedness is not the same in that case anyway.) }
+\itoplassug{Reviewer C, esop, wants to know what happens when there are several stack blocks (see tex for comment).
 We could here emphasize that while we talk about a single stack there could be multiple local read/write-local/execute-capabilities.
@@
 \end{toplas}
-\itoplas{We could further add that this is not the level of abstraction that one would like to reason.
+\itoplassug{We could further add that this is not the level of abstraction that one would like to reason.
 One would like to reason about the programs we actually write - that is programs written in high-level languages.

### 2018-04-25 toplas/paper.tex
Updated figure positions
Added the locality hierarchy and updated text

@@
 
-\begin{figure}[p]
+\begin{figure}[tb]
 \[\arraycolsep=1.4pt
@@
 
-\begin{figure*}
+\begin{figure}[htb]
   \centering
@@
   \label{fig:op-sem}
-\end{figure*}
-
-\begin{wrapfigure}{R}{.3\textwidth}
+\end{figure}
+
+\begin{wrapfigure}{R}{.4\textwidth}
   \centering
@@
     (5) edge (6);
+
+    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
+    \node[main node] (2l) [below of=1l] {$\local$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (1l) edge (2l);
   \end{tikzpicture}
-
-  \caption{Permission hierarchy}
+  \begin{tikzpicture}[main node/.style={}]
+  \end{tikzpicture}
+  \caption{Permission and locality hierarchy.}
   \label{fig:perm-hier}
@@
 of pointers other than capabilities, so we will use the terms interchangeably.
-The available permissions and their ordering are depicted in
-\figurename~\ref{fig:perm-hier}: the permissions include null permission
+The available permissions are null permission
 ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
@@
 ($\rwlx$) and enter ($\entry$), which we will explain below.
-\itoplas{Suggestion: add the locality hierarchy and a short description. Maybe introduce the $\sqsubseteq$ relation here.}
+\itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}. We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
 
@@
 
-\begin{figure}
+\begin{figure}[htb]
   \begin{tikzpicture}[scale=.7, every node={scale=.7}]
@@
 \subsection{Worlds}
-\begin{figure}
+\begin{figure}[htb]
   \centering
@@
   \label{fig:cap-world}
-\end{figure}
+\end{figure}[htb]
 % \dominique{Perhaps we could first try to convey the intuition of a permanent
@@
 \afterpage{
-\begin{figure}[htbp]
+\begin{figure}[htb]
   \centering
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htb]
   \centering

### 2018-04-25 toplas/paper.tex
Moved the macros and added malloc and macro section. Also added a section about reasoning.

@@
 \newcommand{\act}{\var{act}}
+
+\newcommand{\hsframe}{\hs_\var{frame}}
+\newcommand{\size}{\var{size}}
+\newcommand{\rio}{r_{io}}
+\newcommand{\advb}{\var{adv_{base}}}
+\newcommand{\adve}{\var{adv_{end}}}
+\newcommand{\initb}{\var{init}_{base}}
+\newcommand{\inite}{\var{init}_{end}}
+\newcommand{\mrlen}{5cm}
+\newcommand{\retm}{\var{ret}_{\malloc}}
+\newcommand{\reta}{\var{ret}_{\adv}}
+\newcommand{\base}{\var{base}}
+\newcommand{\eend}{\var{end}}
 
@@
 satisfying a certain specification. Malloc and linking tables are described
-further in the next section, but we refer to the technical
-appendix~\citep{technical_appendix} for full details.
 further in the next section.
+\itoplas{The specification of malloc is presented in \sectionname~\ref{sec:malloc-and-macros} as it uses the semantic model we build in Section~\ref{sec:logical-relation}.
+For full details on the linking table, we refer to the technical
+appendix~\citep{technical_appendix}.} \itoplas{(Reviewer A, esop suggest explaining the
+  linking model)\\}
 \itoplassug{Suggestion: We also have the flag table ussed
@@
 \textbf{Calling convention} With the security measures introduced and motivated,
-let us summarize our proposed calling convention: \textit{At program start-up} A
-local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local
-capabilities. \textit{Before returning to the adversary} Clear non-return-value
-registers. Clear the part of the stack we had access to (not just the part we
-used). \textit{Before invoking the adversary} Push activation record to the
+let us summarize our proposed calling convention:
+\begin{description}[font=\normalfont\itshape]
+\item[At program start-up] A
+local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local capabilities.
+\item[Before returning to the adversary] Clear non-return-value
+registers. Clear the part of the stack we had access to (not just the part we used).
+\item[Before invoking the adversary] Push activation record to the
 stack. Create return pointer as local $\entry$-capability to the instructions in
 the record.  Restrict the stack capability to the unused part and clear
-it. Clear non-argument registers. \textit{Before invoking an adversary callback}
-Make sure callback is global. \textit{When invoked by an adversary} Make sure
-received stack pointer has permission $\rwlx$.
-
-\textbf{Reusable macro instructions} 
-\label{para:macros}
-\itoplas{Add examples of macros to this section to make it clear what we mean by macros.}
-\itoplas{Move this subsection forward so it is just before the macros are used, i.e. the example section.}
-\itoplas{Reviewer C, ESOP, asks for high-level specifications for the macros in the style of malloc.
-I am not sure we want to actually include this in the submission, so it is maybe not here we should put our attention.
-(see review in tex comment}
-% Finally, as already mentioned above, the approach relies on `call` and `scall` macro-instructions that are only outlined in the submission, and described by very complicated pseudocode and English prose in the companion technical appendix.
-I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
-We define a number of reusable macros
-capturing the calling convention and other conveniences. All macros that use the
-stack assume a stack pointer in register $r_\stk$. The macro
-\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
-$\var{name}$ from the linking table and stores it in register $r$.  The macros
-\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} add and
-remove elements from the stack. The macro \texttt{\footnotesize{prepstk $r$}} is
-used when a callback is invoked by the adversary and prepares the received stack
-pointer by checking that it has permission $\rwlx$.  The macro
-\texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
-capability in register $r$ in the manner described above. That is, it pushes
-local state (the contents of registers $\overline{r_{\var{priv}}}$) and the
-activation record (return code, return pointer, stack pointer) to the stack,
-creates an $\entry$ return pointer, restricts the stack pointer, clears the
-unused part of the stack, clears the necessary registers and jumps to $r$.  Upon
-return, the private state is restored. The macro \texttt{\footnotesize{mclear
-    $r$}} clears all the memory the capability in register $r$ has authority
-over.  The macro \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the
-registers in $\var{regSet}$. The macro \texttt{\footnotesize{reqglob $r$}}
-checks whether the word in register $r$ is a global capability. The macro
-\texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}} allocates a closure
-where $r$ points to the closure's code and a new environment is allocated (using
-malloc) where the contents of $\overline{r_i}$ is stored. In the code referred
-to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
-
-The technical appendix~\citep{technical_appendix} contains detailed descriptions
-of all the macros.
+it. Clear non-argument registers.
+\item[Before invoking an adversary callback] Make sure callback is global.
+  \item[When invoked by an adversary] Make sure received stack pointer has permission $\rwlx$.
+\end{description}
+
 %\itoplassug{Suggestion: Make it more clear that this CC can be used to protect multiple components (reviewer A, ESOP), see tex comment.\\}
@@
 
+\begin{toplas}
+  \section{Malloc and macros}
+  \label{sec:malloc-and-macros}
+  Assembly code can be difficult for people to easily read and understand as it is close to byte code and offers none of the abstractions of high-level languages.
+In \sectionname~\ref{sec:examples}, we present a number of program examples for the simple capability machine we present in this paper.
+In order to make these examples readable, we introduce a series of macros that each corresponds to a series of actual instructions on the machine.
+
+  In \sectionname~\ref{sec:examples}, we will prove correctness lemmas for each of the program examples.
+Some of the program examples use malloc and with the logical relation defined in \sectionname~\ref{sec:logical-relation}, we can present the specification of malloc.
+
+  In this section, we present the specification of malloc as well as the macros used in \sectionname~\ref{sec:examples}.
+
+\subsection{Malloc specification}
+  \itoplassug{Suggestion: add the malloc specification}
+\begin{definition}[Malloc Specification]
+  \label{spec:malloc}
+  $c_\malloc$ satisfies the specification for malloc iff
+  \[  
+    \begin{array}{l}
+      c_\malloc = ((\entry,\glob),\_,\_,\_) \land\\
+      \exists \iota_{\malloc,0} \ldotp \\
+      \quad \left(\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)])) \right) \wedge \\
+      \quad \iota_{\malloc,0}.v = \perma \; \land \\
+      \quad (\forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
+      \qquad \forall i, n, \size \in \nats \ldotp \forall
+      w_{\var{ret}} \in \Words \ldotp \\
+      \qquad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
+      \qquad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
+      \qquad \quad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
+      \qquad \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
+      \qquad \quad \Rightarrow \\
+      \qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
+      \qquad \qquad \quad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+      \qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
+      \qquad \qquad \qquad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+      \qquad \qquad \qquad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
+      \qquad \qquad \qquad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
+      \qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
+      \qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+      \qquad \qquad \qquad \size - 1 = e'-b' ) \land\\
+      \quad (\forall \Phi \in \ExecConfs \ldotp (\memreg(r_1)
+      \not\in \ints \vee \memreg(r_1) < 0) \land \memreg(\pcreg) = \updatePcPerm{c_\malloc} \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed)
+    \end{array}
+  \]
+\end{definition}
+
+\subsection{Macros}
+\textbf{Reusable macro instructions} 
+\label{para:macros}
+\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
+\itoplassug{Move this subsection forward so it is just before the macros are used, i.e. the example section.\\}
+\itoplassug{Reviewer C, ESOP, asks for high-level specifications for the macros in the style of malloc.
+I am not sure we want to actually include this in the submission, so it is maybe not here we should put our attention.
+(see review in tex comment)\\}
+% Finally, as already mentioned above, the approach relies on `call` and `scall` macro-instructions that are only outlined in the submission, and described by very complicated pseudocode and English prose in the companion technical appendix.
+I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
+We define a number of reusable macros
+capturing the calling convention and other conveniences. All macros that use the
+stack assume a stack pointer in register $r_\stk$. The macro
+\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
+$\var{name}$ from the linking table and stores it in register $r$.  The macros
+\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} add and
+remove elements from the stack. The macro \texttt{\footnotesize{prepstk $r$}} is
+used when a callback is invoked by the adversary and prepares the received stack
+pointer by checking that it has permission $\rwlx$.  The macro
+\texttt{\footnotesize{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
+capability in register $r$ in the manner described above. That is, it pushes
+local state (the contents of registers $\overline{r_{\var{priv}}}$) and the
+activation record (return code, return pointer, stack pointer) to the stack,
+creates an $\entry$ return pointer, restricts the stack pointer, clears the
+unused part of the stack, clears the necessary registers and jumps to $r$.  Upon
+return, the private state is restored. The macro \texttt{\footnotesize{mclear
+    $r$}} clears all the memory the capability in register $r$ has authority
+over.  The macro \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the
+registers in $\var{regSet}$. The macro \texttt{\footnotesize{reqglob $r$}}
+checks whether the word in register $r$ is a global capability. The macro
+\texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}} allocates a closure
+where $r$ points to the closure's code and a new environment is allocated (using
+malloc) where the contents of $\overline{r_i}$ is stored. In the code referred
+to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
+
+The technical appendix~\citep{technical_appendix} contains detailed descriptions
+of all the macros.
+\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
+\end{toplas}
+
+\begin{toplas}
+\section{Reasoning about programs}
+\label{sec:reasoning}
+
+\end{toplas}
+
 \section{Examples}

### 2018-04-25 toplas/paper.tex
Fixed figure and section name issue. typos.

@@
 \renewcommand{\figurename}{Figure}
+\renewcommand{\sectionname}{Section}
 \newcommand{\forcenewline}{$\phantom{v}$\\}
@@
   \begin{itemize}
-  \item Double monotonicity lemma (lemma 77 and 79) with a pointer back to local capability section of popl intro to this section.
+  \item (The lemma is there, consider pointer) Double monotonicity lemma (lemma 77 and 79) with a pointer back to local capability section of popl intro to this section.
   \item A section with the lemmas we use for reasoning about programs
@@
 \begin{toplas}
-\todo[inline]{Look this introductoin and the following section through to check for overlap.
+\todo[inline]{Look this introduction and the following section through to check for overlap.
 Also consider making backwards references to this introduction from the places where we use advanced techniques, so it becomes clear that they are necessary.}
@@
 lemma statements. The examples may look deceivingly short, but it is because
-they use the macro instructions described in Section \ref{para:macros}.
+they use the macro instructions described in \sectionname~\ref{para:macros}.
 The examples would be
@@
 cited, it would be easy to conclude that the returned values are in the value
-relation (see Figure~\ref{fig:logrel}). This gives us a lot of semantic
+relation (see \figurename~\ref{fig:logrel}). This gives us a lot of semantic
 information, like conditions under which they are safe to use and invariants

### 2018-04-25 fully-abstract-comp/ftlr-proof-cases.tex
updated assumptions of FTLR proof

--- 2018-04-13 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \begin{itemize}
-  \item If $\npair{(b,e)}\readCond{\lin}{W}$, then $\npair[n']{(b,e)}\readCond{\lin}{W}$.
+  \item If $\npair{(b,e)} \in \readCond{\lin,W}$, then $\npair[n']{(b,e)} \in \readCond{\lin,W}$.
   \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
@@
   \item If $(w_1,w_2) \in H_\sigma~\sigma~W$, then $(w_1,w_2) \in H_\sigma~\sigma~W'$.
+    \lau{This should hold for any valid $H_\sigma$, but we need to prove it for specific instantiations.}
   \item If $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in
@@
 \begin{proof}
-  Follows easily by inspecting the definitions of $\lrv$, $\readCond$, $\addressable$, $\writeCond$, $\execCond$ and $\xReadCond$ and using Lemma~\ref{lem:purePart-idempotent}.
+  Follows easily by inspecting the definitions of $\lrv$, $\readCond[]{}$, $\addressable{}$, $\writeCond[]{}$, $\execCond[]{}$ and $\xReadCond[]{}$ and using Lemma~\ref{lem:purePart-idempotent}.
 \end{proof}
@@
   Assume
-\[
-  \begin{array}{l}
-  \npair{(\baddr,\eaddr)}\in \readCond{\lin,W}\\
-  \npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}
-  \end{array}
-\]
+  \begin{itemize}
+  \item $\trust = \trusted \vee \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
+  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
+  \end{itemize}
+  and one of the following sets of requirements holds:
+  \begin{enumerate}[label={\roman*)}]
+  \item \begin{itemize}
+    \item $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$
+    \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \trusted$
+    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
+      \begin{itemize}           
+      \item $\src{\Phi}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+      \item $\src{\Phi}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+      \item $\src{\Phi}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+      \item $\src{\Phi}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+      \end{itemize}
+    \end{itemize}
+  \item
+    \begin{itemize}
+    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
+    \item $\trust = \untrusted$
+    \end{itemize}
+  \end{enumerate}
 and show
@@
 \npair{((\rx,\lin),\baddr,\eaddr,\aaddr),
-       ((\rx,\lin),\baddr,\eaddr,\aaddr)} \in \lre(W)
+       ((\rx,\lin),\baddr,\eaddr,\aaddr)} \in \lreg{\trust}(W)
 \]
@@
 \begin{enumerate}
-\item $\npair[n']{[(c,c)]} \in \lrp(W)$ 
-\dominique{Weird that this is an assumption here.
-Shouldn't this be a proof obligation?
-Or rather: I think this should not be here at all but in the value relation instead?
-Remove this from the definition of $\lre$?}
-\lau{$(c,c)$ is not shown or assumed to be in the value relation, so to make sure it respects lineraity, this condition is placed here.
-Further, this does become a proof obligation when you use the FTLR.}
-\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
+\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$ \label{item:reg-ass}
 \item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
@@
 \end{itemize}
+and pick
+\begin{itemize}
+\item $\Phi_S' = \xjumpResult{c}{c}{\Phi_S}$
+\item $\Phi_T' = \xjumpResult{c}{c}{\Phi_T}$
+\end{itemize}
 and show
 \[
-  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
+  \npair[n']{(\Phi_S',\Phi_T')} \in \lro
 \]
-That is, let $i \leq n'$ be given \footnote{ignoring the frame...} and show
-\[
-  \Phi_S \term[i] \Rightarrow \Phi_T \term
-\]
-and 
-\[
-  \Phi_T \term[i] \Rightarrow \Phi_S \term
-\]
-
-{\color{DarkGreen}
-\dominique{Proposed proof refactoring below:}
-First, we prove that one of the following holds:
+In order to do this, we first, we prove that one of the following holds:
 \begin{itemize}
@@
 Now assume that one of the configurations terminates etc.
-
-}
 

### 2018-04-27 toplas/paper.tex
Added some text for the malloc specification section

@@
   \label{sec:malloc-and-macros}
-  Assembly code can be difficult for people to easily read and understand as it is close to byte code and offers none of the abstractions of high-level languages.
-In \sectionname~\ref{sec:examples}, we present a number of program examples for the simple capability machine we present in this paper.
-In order to make these examples readable, we introduce a series of macros that each corresponds to a series of actual instructions on the machine.
-
-  In \sectionname~\ref{sec:examples}, we will prove correctness lemmas for each of the program examples.
-Some of the program examples use malloc and with the logical relation defined in \sectionname~\ref{sec:logical-relation}, we can present the specification of malloc.
-
-  In this section, we present the specification of malloc as well as the macros used in \sectionname~\ref{sec:examples}.
+With the semantic model defined, we would like to show its usefulness by proving
+the correctness of a series of examples that use the calling convention and rely
+on well-bracketedness and local-state encapsulation. However, the programs that
+run on our capability are assembly programs and even program examples that would
+be small in a high-level language, they become big and unintelligible at this
+low level. In order to make our program examples intelligible, we introduce a
+series of low-level abstractions in the form of macros.
+
+In order to make the program examples realistic, we allow them to allocate fresh
+memory by invoking a trusted malloc. Our results are not specific to
+any one implementation of malloc, so rather than providing one implementation,
+we provide a specification that malloc must satisfy.
 
@@
   \itoplassug{Suggestion: add the malloc specification}
+Before we give the malloc specification, we provide some intuition about what
+the malloc specification should look like.
+
+Malloc is assumed to be part of the trusted computing base, as mentioned in
+\sectionname~\ref{sec:capab-mach-with}. A program with access to malloc should
+be able to retain this access even if it invokes code it does not trust. In
+order to achieve this, a capability for malloc must be global. It is also
+important that malloc remains trustworthy to all entities on the machine
+throughout execution. That is, malloc should encapsulate its local state and
+not depend on any external code. To make sure that malloc cannot be tampered
+with through the capability for invoking malloc, we also require the capability
+to have enter permission.
+
+The requirements so far prevent direct tampering in the sense that the
+capability intended for invoking malloc only can be used for invocation. This
+does, however, not prevent the implementation of malloc from colluding with
+other code. Specifically, malloc could depend on the state of a piece of memory
+not just under the control of malloc. We also need to make sure that malloc
+actually behave as intended. That is when malloc is invoked with a length
+argument in the correct register, then it should actually return a capability
+for a fresh piece of memory of that size. It should be fresh in the sense that
+malloc hasn't given out a capability for any part of that memory before and
+won't do it in the future. Finally, we also need to make sure that malloc
+behaves correctly when given a nonsensical length argument such as a negative
+integer or a capability. These requirements relate to the behaviour of malloc,
+so we cannot hope to specify them syntactically. Instead we will use the
+semantic model defined in  Section~\ref{sec:logical-relation} to express this
+part of the specification.
+
+In order to express the semantic requirements on malloc, the specification will
+require the existence of some initial region for malloc. Malloc does not
+disappear from the system, so the malloc region must permanent. This also means
+that the malloc specification should guarantee that use of malloc is safe at any
+point of execution and no matter who has invoked it. To this end, the
+requirement on the initial malloc region should also hold true in any private
+future world.
+
+The implementation of malloc should not depend on any memory outside the
+implementation of malloc. This means that malloc should not have any
+capabilities in its internals that govern other programs. The malloc region's
+memory interpretation function determines what malloc may contain, so we require
+it to allow the same memory segments whether it is with respect to some world
+with the malloc region in it or a world that just contains the malloc region.
+Malloc should retain this property throughout execution, so it must hold true
+for any private future region of the initial malloc region.
+
+The malloc specification also dictates what malloc should do when invoked. We
+specify the two cases where malloc is invoked with a valid and an invalid length
+argument, respectively. When malloc is invoked with an invalid length argument,
+that is a negative integer or a capability, we require malloc to simply fail.
+This part of the specification does not rely on the malloc region as we do not
+need to make sure that the memory invariants are respected. 
+On the other hand,
 \begin{definition}[Malloc Specification]
@@
       \exists \iota_{\malloc,0} \ldotp \\
-      \quad \left(\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)])) \right) \wedge \\
       \quad \iota_{\malloc,0}.v = \perma \; \land \\
-      \quad (\forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
-      \qquad \forall i, n, \size \in \nats \ldotp \forall
+      \quad \left(
+        \begin{array}{l}
+          \forall \iota' \futurestr \iota_{\malloc,0}, W,i \ldotp\\
+          \quad W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
+        \end{array}
+        \right) \wedge \\
+      \quad \left(
+      \begin{array}{l}
+        \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
+        \quad \forall i, n, \size \in \nats \ldotp \forall
         w_{\var{ret}} \in \Words \ldotp \\
-      \qquad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
-      \qquad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
-      \qquad \quad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
-      \qquad \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
-      \qquad \quad \Rightarrow \\
-      \qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-      \qquad \qquad \quad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
-      \qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
-      \qquad \qquad \qquad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-      \qquad \qquad \qquad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
-      \qquad \qquad \qquad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
-      \qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-      \qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-      \qquad \qquad \qquad \size - 1 = e'-b' ) \land\\
-      \quad (\forall \Phi \in \ExecConfs \ldotp (\memreg(r_1)
-      \not\in \ints \vee \memreg(r_1) < 0) \land \memreg(\pcreg) = \updatePcPerm{c_\malloc} \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed)
+        \quad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
+        \quad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
+        \qquad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
+        \qquad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
+        \qquad \Rightarrow \\
+        \qquad \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
+        \qquad \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+        \qquad \qquad \quad \Phi \step[j] \Phi' \land \\
+        \qquad \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+        \qquad \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
+        \qquad \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
+        \qquad \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
+        \qquad \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+        \qquad \qquad \quad \size - 1 = e'-b'
+      \end{array}
+      \right) \land\\
+      \quad \left(
+      \begin{array}{l}
+        \forall \Phi \in \ExecConfs \ldotp\\
+        \quad(\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
+        \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
+        \qquad \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed
+      \end{array}
+        \right)
     \end{array}
@@
 \end{definition}
+
+\itoplassug{Write something about initial capabilities to malloc. We will always
+assume that capabilities for malloc satisfy the malloc specification. The malloc
+specification prevents malloc from giving away internal capabilities and the
+capability machine semantics does not allow capabilities to appear out of thin air.}
+
+\begin{lemma}[Malloc is safe to pass to adversary]
+  \label{lem:malloc-in-vr}
+  If $W(r) \futurestr \iota_{\malloc,0}$, then
+  $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
+\end{lemma}
+
+\itoplassug{Add the assumption that the malloc capability in examples
+  satisfies specification.}
 

### 2018-04-27 local-capability-machine/local-capability-machine.tex
updated execCond in local capability technical report
typos

@@
       & \quad \forall W' \future W \ldotp\\
-      & \qquad\forall a \in [\start,\addrend] \ldotp\\
-      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
+      & \qquad\forall a \in [\start',\addrend'] \subseteq [\start,\addrend] \ldotp\\
+      & \qquad \quad \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\}
     \end{aligned} \\
@@
   Assume $W_2 \futurestr W_1$, $\perm \in \{\exec, \rwx \}$ and $\npair{(\perm,\start,\addrend)} \in \execCond{}(\glob)(W_1)$.
-Now let $W_3 \futurestr W_2$, $\addr \in [\start,\addrend]$, and $n' < n$, and show
-  \[
-    \npair{\stdcap[(\perm,\glob)]} \in \stder(W_3)
+Now let $W_3 \futurestr W_2$, $\addr \in [\start',\addrend'] \subseteq [\start,\addrend]$, and $n' < n$, and show
+  \[
+    \npair{((\perm,\glob),\start',\addrend',\addr)} \in \stder(W_3)
   \]
@@
 \begin{proof}[Proof of Lemma~\ref{lem:execcond-dc}]
-  % TODO
+  Follows easily from definition.
 \end{proof}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:entrycond-mono-pub}]
-  % TODO
+  Follows easily from definition.
 \end{proof}
@@
     Suffices: $\exists W''\futurestr W',\ms_r,\ms'\ldotp$ $\mem' =
-    \ms'\uplus\ms_r\uplus\ms_f$ and $\heapSat[\ms']{n-i}{W''}$\\
+    \ms'\uplus\ms_r\uplus\ms_f$ and $\heapSat[\ms']{n''-i}{W''}$\\
     By: definition of $\observations(W')$
@@
         r}P(r')\right)\uplus\ms_f$. Then $\ms\uplus\ms_f = P(r) \uplus
-    \ms_{\var{frame}}$ and $\heapSat[P(r)]{n''}{W'(r).H~(W'(r).s)~(\xi^{-1}(W'))}$.
+    \ms_{\var{frame}}$ and $\npair[n'']{P(r)} \in W'(r).H~(W'(r).s)~(\xi^{-1}(W'))$.
     Easy from the previous point.
-  \item $\heapSat[P(r)]{n''}{W'(r).H~(W'(r).s)~(\xi^{-1}([r\mapsto W'(r)]))}$, i.e.
+  \item $\npair[n'']{P(r)} \in W'(r).H~(W'(r).s)~(\xi^{-1}([r\mapsto W'(r)]))$, i.e.
     $\heapSat[P(r)]{n''}{[r\mapsto W'(r)]}$.\\
@@
       $\npair[n'']{P(r)} \in H(s)(\xi^{-1}(W'))$) and $W'' \futurewk W'$ and the
-      public monotonicity and downwards closedness of all islands, and finally
+      public monotonicity and downwards closedness of all regions, and finally
       the definition of $W''$.
@@
       By Lemma~\ref{lem:stdvr-mono-wk}, the fact that $W''\futurewk W'$,
-      Lemma~\ref{lem:stdvr-dc}, and teh fact that
+      Lemma~\ref{lem:stdvr-dc}, and the fact that
       $\npair[n'']{w_{\var{ret}}}\in\stdvr(W')$, which follows from
@@
       By: definition of $\stder(W'')$, using the previous point and the facts
-      that
+      that\\
       $\npair[n''-j]{\memreg[\Phi]\update{r_1}{((\rwx,\glob),b',e',b')}}\in\stdrr(W'')$,
@@
       $(\reg\update{\pcreg}{((\exec,\glob),\start,\addrend,\addr)},\ms\uplus\ms_f)
-      \step[j] \Phi'$ using Lemma~\label{lem:determinacy}.
+      \step[j] \Phi'$ using Lemma~\ref{lem:determinacy}.
     \item $\exists W''' \futurestr W'', \hs_r, \hs'' \ldotp$ $\heap' = \hs'' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs'']{n-i}{W'''}$.\\

### 2018-04-27 toplas/paper.tex
small changes and addition to toplas paper

@@
 
-For reasons of space, some details and all proofs have been omitted; please
-refer to the technical appendix~\citep{technical_appendix} for
-those.\dominique{upload TA as supplementary material instead?}
+We have written a technical appendix~\citep{technical_appendix} which contains
+additional details and proofs left out from this paper.
 
@@
 load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and arithmetic addition operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and
+and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and
 $\minus{}{}{}$, operating only on numbers). There are three instructions for
@@
 In general, should we add all the missing definitions and make this document closer to being self-contained.}
-shows an excerpt of the operational semantics for a
+shows the operational semantics for a
 few representative instructions. Essentially, a configuration $\Phi$ either
@@
 figure shows for instructions $i$ the result of executing them in configuration
-$\Phi$. $\fail$ and $\halt$ obviously fail and halt respectively. $\move{}{}$
+$\Phi$. The instructions $\fail$ and $\halt$ obviously fail and halt respectively. $\move{}{}$
 simply modifies the register file as requested and updates the $\pcreg$ to the
@@
   for assertions, maybe we should add a description of it here.\\}
-\itoplassug{Reviewer C, popl, would have liked a discussion of the relationship
-  between this capability machines and old capability machines from the
-  litterature (Cambridge CAP \& IBM System 39, see tex comment).\\}
+% \itoplassug{Reviewer C, popl, would have liked a discussion of the relationship
+%   between this capability machines and old capability machines from the
+%   litterature (Cambridge CAP \& IBM System 39, see tex comment).\\}
  % * I'm old, so I'd appreciate a brief discussion LCM's relationship
@@
   \item (The lemma is there, consider pointer) Double monotonicity lemma (lemma 77 and 79) with a pointer back to local capability section of popl intro to this section.
-  \item A section with the lemmas we use for reasoning about programs
-    \begin{itemize}
-    \item lemma 58, scall works. Include description of how it handles the common things in call and leaves the argumentation for callee code and continuation.
-    \item lemma 59, malloc works. If this is included, we should probably also include the malloc specification.
-    \item lemma 60, create closure works.
-    \item Include a description about how the FTLR is used to reason about unknown code.
-    \item Mention that this is a step on the way towards having a program logic, and then in the discussion discuss why a program logic for this language is not what we need.
-    \end{itemize}
   \item Reviewer B, ESOP suggests inferring general security properties from the LR.
@@
   \label{fig:cap-world}
-\end{figure}[htb]
+\end{figure}
 % \dominique{Perhaps we could first try to convey the intuition of a permanent
@@
 change over time. In other words, they are protocols for the region's
-memory. These are similar to what has been used in logical relations for
-high-level
-languages~\citep{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. Protocols
+memory. Protocols
 imposed by permanent regions stay in place indefinitely. Any capability, local
@@
       &\stdvr(W)\defeq 
-        \begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \} \union \{ \npair{\stdcap[(\noperm,\gl)] } \}
+        \begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \} \union \\
+          &
+          \{ \npair{\stdcap[(\noperm,\gl)] } \}
         \union \\
-        % & \left\{
-        %     \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-        %   \right\}
-        % \union \\
+        & \left\{
+            \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
+          \right\}
+        \union \\
         & \left\{
@@
         \union \\ 
-        % & \left\{
-        %   \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
-        %   \begin{array}{l}
-        %     \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-        %     \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
-        %     \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-        %   \end{array}
-        %   \right\}
-        % \union \\
+        & \left\{
+          \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
+          \begin{array}{l}
+            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
+            \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
+            \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
+          \end{array}
+          \right\}
+        \union \\
         & \left\{
@@
             \end{array}
-          \right\} \\
-          \union
-        & \dots \small{\text{\itshape and so on for permissions $\readonly$, $\readwritel$, $\exec$, and $\rwx$.}}
+          \right\}
       \end{aligned}
@@
 \label{sec:reasoning}
+\itoplassug{
+  \begin{itemize}
+  \item A section with the lemmas we use for reasoning about programs
+    \begin{itemize}
+    \item lemma 58, scall works. Include description of how it handles the
+      common things in call and leaves the argumentation for callee code and
+      continuation.
+    \item lemma 59, malloc works. If this is included, we should probably also
+      include the malloc specification.
+    \item lemma 60, create closure works.
+    \item Include a description about how the FTLR is used to reason about
+      unknown code.
+    \item Mention that this is a step on the way towards having a program logic,
+      and then in the discussion discuss why a program logic for this language
+      is not what we need.
+    \end{itemize}
+  \end{itemize}
+}
 

### 2018-04-27 toplas/paper.tex
Added a word on modularity to the section on stack and return pointer management.

@@
 \begin{toplas}
-  \textbf{Modularity} The calling convention ensures well-bracketed calls and local-state encapsulation for the caller, but not the callee.
-In the above presentation to make it easy to distinguish between the parties involved, we present the callee as some adversarial code that we do not trust.
-In reality, the callee could be well-behaved and wish to ensure well-bracketed calls and local-state encapsulation.
-The calling convention puts no restriction on the callee that the caller itself does not follow, so by following the calling convention, the callee would also get well-bracketedness and local-state encapsulation.
-Generally, the calling convention ensures the two properties for anyone that employ it, so it scales to the setting where multiple distrusting parties need to invoke each other and each of them can protect themself by using the calling convention.
+  \textbf{Modularity} The calling convention ensures well-bracketed calls and
+  local-state encapsulation for the caller, but not the callee. In the above
+  presentation to make it easy to distinguish between the parties involved, we
+  present the callee as some adversarial code that we do not trust. In reality,
+  the callee could be well-behaved and wish to ensure well-bracketed calls and
+  local-state encapsulation. The calling convention puts no restriction on the
+  callee that the caller itself does not follow, so by following the calling
+  convention, the callee would also get well-bracketedness and local-state
+  encapsulation. Generally, the calling convention ensures the two properties
+  for anyone who employs it, so it scales to the setting where multiple
+  distrusting parties need to invoke each other and each of them can protect
+  themself by using the calling convention.
 \end{toplas}

### 2018-04-28 toplas/paper.tex
Added definition of standard regions to the paper

@@
 \end{figure}
+\begin{figure}[htb]
+  \begin{align*}
+  H^\nwl_A,  H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
+  \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
+  \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
+  H^\pwl_A \; s \; \hat{W} \defeq{} & \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = A \land \\
+      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms} \} \\
+  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
+  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
+  H^\nwl_A \; s \;\hat{W} \defeq{} & \left\{\npair{\hs} \middle|
+    \begin{aligned}
+      &\dom(\hs) = A \land \\
+      &\forall \addr \in A \ldotp \\
+      &\quad \nonlocal{\ms(\addr)} \land\\ 
+      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms} \}
+\end{align*}
+
+  \caption{The standard permit write-local and no write-local regions.}
+  \label{fig:std-reg}
+\end{figure}
 }
+
 %subsection introduction
@@
 imposes safety as a requirement on the values contained. This safety requirement
-is formulated in terms of a standard region $\iota^\pwl_{\start,\addrend}$. The
-definition of that standard region is omitted for space reasons, but it simply
+is formulated in terms of a standard region $\iota^\pwl_{[\start,\addrend]}$. It simply
 requires all the words in the range $[\start,\addrend]$ to be safe, i.e.\ in the
-value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{\start,\addrend}$
+value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{[\start,\addrend]}$
 means that $W(r)$ must accept only safe values like
-$\iota^\pwl_{\start,\addrend}$, but can be even more restrictive if desired. The
+$\iota^\pwl_{[\start,\addrend]}$, but can be even more restrictive if desired. The
 read condition also takes into account the locality of the capability because,
@@
 all active (non-revoked) regions when the locality $\gl$ is local, but only the
-permanent regions when $\gl$ is global. The 
-the definition of the standard region
-$\iota^\pwl_{\start,\addrend}$ can be found
-in~\citep{technical_appendix}; it makes use of the isomorphism from
+permanent regions when $\gl$ is global. The definition of the standard region
+$\iota^\pwl_{[\start,\addrend]}$ can be found
+in \figurename~\ref{fig:std-reg}; it makes use of the isomorphism from
 Theorem~\ref{thm:world-existence}.
@@
 region. For the write-local capabilities the write condition is applied with the
-standard region $\iota^\pwl_{\start,\addrend}$ that we described previously. For
+standard region $\iota^\pwl_{[\start,\addrend]}$ that we described previously. For
 the write capabilities we use a different standard region
-$\iota^\nwl_{\start,\addrend}$ which requires that the words in
+$\iota^\nwl_{[\start,\addrend]}$ which requires that the words in
 $[\start,\addrend]$ are non-local and safe. As before, we use
@@
 local capabilities.
+The standard region $\iota^\nwl_{[\start,\addrend]}$ is defined in \figurename~\ref{fig:std-reg}.
 

### 2018-04-28 toplas/paper.tex
Added the following definitions to appendix in paper: nsubsim, erasure, and acvtiveReg
Added implementation of scall figure and macros used by scall to appendix

@@
 \itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
+
+\begin{figure}[htb]
+  \centering
+\begin{lstlisting}
+// push private registers to the stack
+  push $r_{\var{priv},1}$
+  ...
+  push $r_{\var{priv},n}$
+// push restoration code to the stack
+  push $\encode(i_1)$
+  ...
+  push $\encode(i_4)$
+// push old pc to the stack
+  move r_t1 pc
+  lea r_t1 $\var{ofc}$ 
+  push r_t1
+// push stack pointer to the stack
+  push r_stk
+// set up protected return pointer
+  move r_0 r_stk
+  lea r_0 $\var{ofc_\var{act}}$ // -5 is the offset to the first instruction of the activation code
+  restrict r_0 $\encodePermPair((\local,\entry))$
+// restrict stack capability to unused part
+  geta r_t1 r_stk
+  plus r_t1 r_t1 1
+  getb r_t2 r_stk
+  subseg r_stk r_t1 r_t2
+// clear unused part of the stack
+  mclear r_stk
+// clear non-argument registers
+  rclear $R$
+  jmp r
+after:
+// pop the restore code
+  pop r_t1
+  pop r_t1
+  pop r_t1
+  pop r_t1
+// pop the private state into approriate registers
+  pop $r_{\var{priv},1}$
+  ...
+  pop $r_{\var{priv},n}$
+\end{lstlisting}
+  \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
+    r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The restoration
+  code is given in \figurename~\ref{fig:scall-rest-code}. The variable
+  $\var{ofc}$ is the offset to the label after, and $\var{ofc_\var{act}} =
+  -5$ which is the offset to the first instruction of the activation record. The
+set $R = \RegName \setminus
+\{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
+  \label{fig:scall-impl}
+\end{figure}
+\begin{figure}[htb]
+  \centering
+  \begin{lstlisting}
+ $i_1$ = move r_t1 pc
+ $i_2$ = lea r_t1 $\var{ofc}$
+ $i_3$ = load r_stk r_t1
+ $i_4$ = pop pc
+\end{lstlisting}
+  \caption{The restoration code used in \texttt{scall}. The variable $\var{ofc}$ is 5
+    which is the offset to the address where the old stack pointer is stored on
+    the stack.} 
+  \label{fig:scall-rest-code}
+\end{figure}
+
 \end{toplas}
@@
 %% Appendix
-% \appendix
-% \section{Appendix}
-
+\appendix
+\section{Appendix}
+\subsection{Logical relation}
+\label{app:logical-relation}
 % Text of appendix \ldots
+\textbf{$n$-subset simulation}
+\begin{mathpar}
+  \inferrule{(s,\phi_\pub,\phi) = (s',\phi_\pub',\phi') \\
+    \forall \hat{W} \ldotp H \; s \; \hat{W} \nsubeq H' \; s' \; \hat{W} }
+  { (v,s,\phi_\pub,\phi,H) \nsubsim (v',s',\phi_\pub',\phi',H')}
+\end{mathpar}
+\textbf{Erasure}
+\begin{align*}
+  \lfloor W \rfloor_S \defeq \lambda r \ldotp 
+  \begin{cases}
+    W(r) & W(r).v \in S\\
+    \bot & \text{otherwise}
+  \end{cases}
+\end{align*}
+\textbf{Active region projection}
+\begin{align*}
+  \activeReg{} & : \Worlds \fun 2^\RegionNames \\
+  \activeReg{W} & \defeq \dom(\erase{W}{\perma,\temp})
+\end{align*}
+
+\subsection{Macros}
+Implementation of the macros used in \texttt{scall}. Implementations of the
+macros not presented here can be found in the technical
+appendix~\citep{technical_appendix}.\\
+\label{app:macros}
+\texttt{push $r$}
+\begin{lstlisting}
+  lea r_stk 1
+  store r_stk r
+\end{lstlisting}
+\texttt{pop $r$}
+\begin{lstlisting}
+  load r r_stk
+  minus r_t1 0 1
+  lea r_stk r_t1
+\end{lstlisting}
+\texttt{rclear $r_1,\dots, r_n$}
+\begin{lstlisting}
+move $r_1$ 0
+move $r_2$ 0
+$\dots$
+move $r_n$ 0
+\end{lstlisting}
+\texttt{mclear $r$}
+\begin{lstlisting}
+        move r_t $r$
+        getb r_t1 r_t
+        geta r_t2 r_t
+        minus r_t2 r_t1 r_t2
+        lea r_t r_t2
+        gete r_t2
+        minus r_t1 r_t2 r_t1
+        plus r_t1 r_t1 1
+        move r_t2 pc
+        lea r_t2 $\var{ofc}_\var{end}$ 
+        move r_t3 pc
+        lea r_t3 $\var{ofc}_\var{iter}$
+iter:
+        jnz r_t2 r_t1
+        store r_t 0
+        lea r_t 1
+        plus r_t1 r_t1 1
+        jmp r_t3
+end:
+        move r_t 0
+        move r_t1 0
+        move r_t2 0
+        move r_t3 0
+\end{lstlisting}
+Where $\var{ofc}_\var{end}$ and $\var{ofc}_\var{iter}$ are the offsets to the
+label \texttt{end} % 9?
+and \texttt{iter}, % 2?
+respectively.
 

### 2018-04-28 toplas/paper.tex
Added cofe definitions to appendix

@@
 
+\subsection{Complete ordered family of equivalences (c.o.f.e)}
+\label{app:cofe}
+This is an excerpt from \citet{Birkedal:tutorial-notes} about c.o.f.e.'s.
+\begin{definition}[o.f.e.] An \emph{ordered family of equivalence} (o.f.e) is a
+  set and a family of equivalences $\left(X, \left( \nequal \right)_{n=0}^\infty
+  \right)$ that satisfy the following properties: 
+  \begin{itemize}
+  \item $\nequal[0]$ is the total relation on $X$
+  \item $\forall n \ldotp \forall x,y \in X \ldotp x \nequal[n+1] y \Rightarrow
+    x \nequal y$
+  \item $\forall x,y \in X \ldotp (\forall n\ldotp x \nequal y) \Rightarrow x = y$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}[c.o.f.e.] A \emph{complete} ordered family of equivalences
+  (c.o.f.e) is an o.f.e $\left(X, \left( \nequal \right)_{n=0}^\infty \right)$
+  where all Cauchy sequences have a limit.  
+\end{definition}
+
 \subsection{Macros}

### 2018-04-30 toplas/paper.tex
Added an FTLR proof sketch.

@@
 \end{theorem}
+\begin{toplas}
+  \begin{proof}
+    Induction over $n$. By definition of $\stder(W)$, show
+    $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
+    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg}$, and
+    $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
+    $\mem'$, and $i \leq n'$ be given and assume $\Phi \step[i]
+    (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
+    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halting
+    configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
+    \var{conf}' \step[i-1] (\halted,\mem')$.
+
+    The rest of the proof roughly consists of two parts (1) argue that
+    $\var{conf}'$ is consistent with the world in the sense that the
+    register-file and memory still respect the world. (2) Argue that the rest of
+    the execution respects the world. The second case, relies on where the
+    $\pcreg$ in $\var{conf}'$ comes from. If the step to $\var{conf}'$ was a
+    jump, then use the fact that the register-file contains safe values in the
+    initial configuration. On the other hand, if the $\pcreg$ was just
+    incremented, then we need to apply the induction hypothesis which is
+    possible as all of the permission based conditions are downwards closed.
+    For instance in the case where the $\pcreg$ is read-execute, i.e.\ $\perm =
+    \exec$, we also assume $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
+    which entails $\npair[n'-1]{(\start,\addrend)} \in \readCond{}(\gl)(W)$ and
+    allows us to use the induction hypothesis.
+    In order to argue (1), we just need to consider the configurations the
+    operational semantics allows us to get to from the initial state. If we
+    consider the memory in $\var{conf}'$, then it either (a) remains unchanged
+    or (b) one address has been updated and in that case, the register-file contains an 
+    appropriate capability for writing. The latter case, (b) corresponds to the first
+    step executing a $\texttt{store}$ instruction, and the former (a) corresponds to
+    any other instruction.
+    In case (a), we can conclude that the memory still respects the world just
+    by downwards closure of memory satisfaction. In the other case, (b), we use
+    an auxiliary lemma called \emph{Conditions for \texttt{store} are
+      sufficient} (see the appendix, Lemma~\ref{lem:conds-store-stuff})
+    that states that given all the assumptions we have done so far, the updated
+    memory still satisfies the world.
+    All the possible changes to the register-file correspond to the effect of
+    some instruction, so for the non-trivial instructions we use a sufficiency
+    lemma.
+
+    The complete proof can be found in the technical appendix~\citep{technical_appendix}.
+  \end{proof}
+\end{toplas}
 The permission-based conditions of Theorem~\ref{thm:ftlr} make sure that the
@@
 examples depend on these properties for correctness.
-See the technical appendix~\citep{technical_appendix} for a detailed
-proof (by induction over the step-index $n$) of the theorem.
+% See the technical appendix~\citep{technical_appendix} for a detailed
+% proof (by induction over the step-index $n$) of the theorem.
+
 % LB: following omitted for brevity:
@@
 
+\subsection{Sufficiency lemma}
+ \begin{lemma}[Conditions for store instruction are sufficient]
+   \label{lem:conds-store-suff}
+   If 
+   \begin{itemize}
+   \item $\ms = \ms' \uplus \ms_f$
+   \item $\heapSat[\ms']{n}{W}$
+   \item $((\perm,\gl),\start,\addrend,\addr) = c$
+   \item $\npair{c}\in\stdvr(W)$
+   \item $\writeAllowed{\perm}$
+   \item $\withinBounds{\var{c}}$
+   \item $\npair{\var{w}}\in\stdvr(W)$
+   \item if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in
+     \{\rwlx,\readwritel \}$
+   \end{itemize}
+
+   then $\addr \in \dom(\ms')$ (i.e. $\ms\update{a}{w} =
+   \ms'\update{a}{w}\uplus\ms_f$) and
+   $\heapSat[{\ms'\update{\addr}{\var{w}}}]{n}{W}$
+ \end{lemma}
+
 \end{document}

### 2018-05-01 toplas/paper.tex
Wrote the rest of the malloc section text

@@
 \subsection{Worlds}
+\label{subsec:worlds}
 \begin{figure}[htb]
@@
 With the semantic model defined, we would like to show its usefulness by proving
-the correctness of a series of examples that use the calling convention and rely
-on well-bracketedness and local-state encapsulation. However, the programs that
+the correctness of a series of examples that rely
+on well-bracketedness and local-state encapsulation and use the calling
+convention to enforce these properties. However, the programs that
 run on our capability are assembly programs and even program examples that would
@@
 
-In order to make the program examples realistic, we allow them to allocate fresh
-memory by invoking a trusted malloc. Our results are not specific to
-any one implementation of malloc, so rather than providing one implementation,
+To make the program examples realistic, they have access to a trusted malloc, so
+they can allocate new memory. Our results are not specific to
+any one implementation of malloc, so rather than providing the implementation,
 we provide a specification that malloc must satisfy.
@@
   \itoplassug{Suggestion: add the malloc specification}
-Before we give the malloc specification, we provide some intuition about what
-the malloc specification should look like.
+Before we give the malloc specification, we provide some intuition about its requirements.
 
@@
 not depend on any external code. To make sure that malloc cannot be tampered
-with through the capability for invoking malloc, we also require the capability
-to have enter permission.
-
-The requirements so far prevent direct tampering in the sense that the
-capability intended for invoking malloc only can be used for invocation. This
-does, however, not prevent the implementation of malloc from colluding with
-other code. Specifically, malloc could depend on the state of a piece of memory
-not just under the control of malloc. We also need to make sure that malloc
-actually behave as intended. That is when malloc is invoked with a length
-argument in the correct register, then it should actually return a capability
-for a fresh piece of memory of that size. It should be fresh in the sense that
-malloc hasn't given out a capability for any part of that memory before and
-won't do it in the future. Finally, we also need to make sure that malloc
-behaves correctly when given a nonsensical length argument such as a negative
-integer or a capability. These requirements relate to the behaviour of malloc,
-so we cannot hope to specify them syntactically. Instead we will use the
-semantic model defined in  Section~\ref{sec:logical-relation} to express this
-part of the specification.
+with directly through the malloc capability, we require it to have enter
+permission.
+
+The requirements so far are syntactic and prevent direct tampering in the sense
+that the capability intended for invoking malloc only can be used for
+invocation. This does, however, not prevent the implementation of malloc from
+colluding with other code. Specifically, malloc could depend on the state of a
+piece of memory not just under the control of malloc. We also need to make sure
+that malloc actually behave as intended. That is when malloc is invoked with a
+length argument, then it actually returns a capability for a fresh piece of
+memory of that size. It should be fresh in the sense that malloc hasn't given
+out a capability for any part of that memory before and won't do it in the
+future. Finally, we also need to make sure that malloc behaves correctly when
+given a nonsensical length argument such as a negative integer or a capability.
+These requirements relate to the behaviour of malloc, so we cannot hope to
+express them syntactically. Instead we will use the semantic model defined in
+Section~\ref{sec:logical-relation} to express this part of the specification.
 
@@
 require the existence of some initial region for malloc. Malloc does not
-disappear from the system, so the malloc region must permanent. This also means
-that the malloc specification should guarantee that use of malloc is safe at any
-point of execution and no matter who has invoked it. To this end, the
+disappear from the system, so the malloc region must be permanent. This also
+means that the malloc specification should guarantee that use of malloc is safe
+at any point of execution no matter who has invoked it. To this end, the
 requirement on the initial malloc region should also hold true in any private
 future world.
+% priv future means that malloc should not depend on local capabilities?
 
@@
 implementation of malloc. This means that malloc should not have any
-capabilities in its internals that govern other programs. The malloc region's
-memory interpretation function determines what malloc may contain, so we require
-it to allow the same memory segments whether it is with respect to some world
-with the malloc region in it or a world that just contains the malloc region.
-Malloc should retain this property throughout execution, so it must hold true
-for any private future region of the initial malloc region.
-
-The malloc specification also dictates what malloc should do when invoked. We
-specify the two cases where malloc is invoked with a valid and an invalid length
-argument, respectively. When malloc is invoked with an invalid length argument,
-that is a negative integer or a capability, we require malloc to simply fail.
-This part of the specification does not rely on the malloc region as we do not
-need to make sure that the memory invariants are respected. 
-On the other hand,
+capabilities in its internals that govern other programs. As explained in
+\sectionname~\ref{subsec:worlds}, a region's state interpretation function
+determines what memory segments it permits with respect to some world. The state
+interpretation function of malloc should permit the same regions with respect to
+a world with the malloc region as it would with respect to a world that only
+contains the malloc region. In other words, the remaining regions in the world
+do not influence which regions malloc accepts. Malloc should retain this
+property throughout execution, so it must hold true for any private future
+region of the initial malloc region.
+
+The malloc specification will also dictate what malloc should do when invoked.
+Malloc can be invoked with either a valid or an invalid length argument. In the
+case where malloc is invoked with an invalid argument, that is a negative
+integer or a capability, we require malloc to simply fail. This part of the
+specification does not rely on the malloc region as failing is considered
+safe behaviour. When malloc is invoked with a valid length, then malloc should
+return a fresh memory segment of the correct length.
+\todo[inline]{Everything below this is work in progress..}
+Malloc may have been invoked a number of times before any given invocation, so
+the internal data structures of malloc may be in a different state than it
+started in.
+We know that the memory of malloc is in a state consistent with the
+protocol dictated by the initial malloc region in the sense that we can assume
+that malloc is invoked in a configuration where part of it satisfies a private
+future region of the initial malloc region. Knowing the state of the malloc
+region allows us to express what we mean by a \emph{fresh} region. This means
+that malloc should not keep any capabilities for the freshly allocated region.
+To ensure this, we require the footprint of malloc after the call to be disjoint
+from the freshly allocated memory. Further, the footprint should satisfy a
+public future region of the malloc region the invocation happened in. This means
+that it cannot depend on the freshly allocated region. It further allows
+subsequent calls to malloc as it provides one of the requirements in the malloc
+specification.
+
+Malloc is part of the trusted computing base. If we cannot trust malloc, then we
+cannot use the memory it allocates as we have no idea whether it is aliased by
+some untrusted program. Because of this assumption, we also require malloc to
+promise that the non-argument registers of the register-file remain unchanged
+after the call. This allows the caller to keep private capabilities in the
+register file. This is strictly speaking not necessary as we can use our calling
+convention to protect our local state.
+
 \begin{definition}[Malloc Specification]
@@
 \end{definition}
-
-\itoplassug{Write something about initial capabilities to malloc. We will always
-assume that capabilities for malloc satisfy the malloc specification. The malloc
-specification prevents malloc from giving away internal capabilities and the
-capability machine semantics does not allow capabilities to appear out of thin air.}
-
+As described previously, the specification of malloc ensure that malloc has no
+capabilities pointing \emph{out} of malloc. It does, however, not say anything about
+capabilities that point \emph{in} to malloc. We obviously can obviously not
+allow this if we want to trust malloc. We have chosen to keep the malloc
+specification simple and let this assumption be in the lemmas that use malloc.
+It is sufficient for these lemmas to require that there are no outside
+capabilities for malloc in the initial configuration as capabilities cannot
+appear out of thin air and the malloc specification makes sure that capabilities
+are not leaked.
+
+Malloc should not just be available to trusted programs, but also to possibly
+malicious programs. The specification should therefore entail that the malloc
+capability is safe which would make it safe to let even malicious programs have
+access to it.
 \begin{lemma}[Malloc is safe to pass to adversary]
   \label{lem:malloc-in-vr}
-  If $W(r) \futurestr \iota_{\malloc,0}$, then
+  For $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
@@
 state.  On the other hand, \texttt{\footnotesize{f2}} uses malloc to
-allocate memory for its local state and uses an activation-record
-based calling convention (described in the technical appendix) to run
+allocate memory for its local state and uses a calling convention based on heap
+allocated activation records (described in the technical appendix) to invoke
 the adversarial code.

### 2018-05-01 toplas/paper.tex
Added malloc specification assumption to the first example lemma
Added the full correctness lemmas to the appendix

@@
   \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc with
+    $\iota_{\malloc,0}$
+  \item $\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}$
   \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
@@
 
+ \subsection{Example correctness lemmas}
+\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
+  \label{lem:correctness-f1-app}
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    m & \defeq \hs_{f1} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus 
+        \ms_{\malloc} \uplus 
+        \hs_{\var{frame}} 
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
+  \end{itemize}
+  where 
+  \begin{align*}
+    &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link+1]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
+  \item $\ms_\flag = [\flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+  \item $\hs_\adv$ contains a global read-only capability for $\hs_\link$ on its first address. The remaining cells of the memory segment only contain instructions.
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f1}},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
+\begin{lemma}[Correctness lemma for \texttt{f2}]
+  \label{lem:correctness-f2-app}
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f2} & \defeq ((\rwx,\glob),\mathtt{f2}-\olf,\mathtt{2f},\mathtt{f2}) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
+    \reg & \in \Regs \\
+    m & \defeq \hs_{f2} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus 
+        \ms_{\malloc} \uplus 
+        \ms_{\var{stk}} \uplus
+        \ms_{\var{frame}} 
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
+  \end{itemize}
+  where 
+  \begin{align*}
+    &\dom(\hs_{f2}) = [\mathtt{f2}-\olf,\mathtt{2f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link+1]\\
+    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $\ms_{f2}(\mathtt{f2}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f2}(\mathtt{f2}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f2}$ contains the code of $f2$.
+  \item $\ms_\flag = [\flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+  \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f2}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+\begin{lemma}[Correctness lemma for \texttt{f3}]
+  \label{lem:correctness-f3}
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{f3} & \defeq ((\rwx,\glob),\mathtt{f3}-\olf,\mathtt{3f},\mathtt{f3}) \\
+    c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link) \\
+    \reg & \in \Regs \\
+    m & \defeq \hs_{f3} \uplus 
+        \hs_\flag \uplus                
+        \ms_{\var{link}} \uplus 
+        \hs_\adv \uplus 
+        \ms_{\malloc} \uplus 
+        \ms_{\var{stk}} \uplus
+        \ms_{\var{frame}} 
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc.
+  \end{itemize}
+  where 
+  \begin{align*}
+    &\dom(\hs_{f3}) = [\mathtt{f3}-\olf,\mathtt{3f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link+1]\\
+    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $\ms_{f3}(\mathtt{f3}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f3}(\mathtt{f3}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f3}$ contains the code of $f3$.
+  \item $\ms_\flag = [\flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
+  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
+  \end{itemize}
+  if 
+  \[
+    (\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
+ \begin{lemma}[Correctness of $g1$]
+  \label{lem:correctness-g1}
+  For all $n \in \nats$
+  let
+  \begin{align*}
+    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
+    c_{g1} & \defeq ((\entry,\glob),\mathtt{g1}-\olf,\mathtt{4f},\mathtt{g1}) \\
+    c_\stk & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
+    c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
+    c_\link & \defeq ((\readonly,\glob),\link,\link,\link) \\
+    m & \defeq \hs_{g1} \uplus 
+        \ms_\flag \uplus                
+        \ms_\link \uplus 
+        \ms_\adv \uplus 
+        \ms_\malloc \uplus 
+        \ms_\stk \uplus
+        \ms_{\var{frame}} 
+  \end{align*}
+  where 
+  \begin{itemize}
+  \item $c_\malloc$ satisfies the specification for malloc with $\iota_{\malloc,0}$
+  \end{itemize}
+  \begin{align*}
+    &\dom(\hs_{g1}) = [\mathtt{g1}-\olf,\mathtt{4f}] \\
+    &\dom(\hs_\flag) = [\flag,\flag] \\
+    &\dom(\ms_\link) = [\link,\link]\\
+    &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
+    &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
+    &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}
+  \end{align*}
+  and
+  \begin{itemize}
+  \item $\ms_{g1}(\mathtt{g1}-\olf) = ((\readonly,\glob),\link,\link,\link)$, $\ms_{g1}(\mathtt{g1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{g1}$ contains the code of $g1$ immediately followed by the code of $f4$.
+  \item $\ms_\flag = [\flag \mapsto 0]$
+  \item $\ms_{\var{link}} = [\link \mapsto c_\malloc]$
+  \item $\hs_\adv(\start_\adv) = c_\link$ and all other addresses of $\ms_\adv$ contain instructions.
+  \item $\forall a \in \dom(\ms_\stk) \ldotp \ms_\stk(a) = 0$ %This condition could be weakened to \memSat{\ms_\stk}{[\iota^\pwl(\dom(\ms_\stk))]}
+  \end{itemize}
+  if 
+  \[
+    (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step[n] (\halted,m'),
+  \]
+  then
+  \[
+    m'(\flag) = 0
+  \]  
+\end{lemma}
+
 \end{document}

### 2018-05-02 toplas/paper.tex
Minor note and move a TODO to a more appropriate place.

@@
 
+\dominique{I think we should include a comparison with the conference paper, and explain what the delta is for this journal version.}
 In this paper, we make two main contributions: (1) an alternative calling
@@
 % p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
-\itoplassug{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}
 \itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
@@
 to what we provide for local capabilities.
+
+\itoplassug{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}
 

### 2018-05-02 toplas/paper.tex
Wrote a page for the macros section

--- 2018-05-01 toplas/paper.tex
@@
 
-\subsection{Macros}
-\textbf{Reusable macro instructions} 
-\label{para:macros}
-\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
-\itoplassug{Move this subsection forward so it is just before the macros are used, i.e. the example section.\\}
-\itoplassug{Reviewer C, ESOP, asks for high-level specifications for the macros in the style of malloc.
+\subsection{Reusable macro instructions}
+\label{subsec:macros}
+%\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
+%\itoplassug{Reviewer C, ESOP, asks for high-level specifications for the macros in the style of malloc.
 I am not sure we want to actually include this in the submission, so it is maybe not here we should put our attention.
@@
 I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
-We define a number of reusable macros
-capturing the calling convention and other conveniences. All macros that use the
-stack assume a stack pointer in register $r_\stk$. The macro
-\texttt{\footnotesize{fetch $r$ $\var{name}$}} fetches the capability related to
-$\var{name}$ from the linking table and stores it in register $r$.  The macros
-\texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}} add and
-remove elements from the stack. The macro \texttt{\footnotesize{prepstk $r$}} is
-used when a callback is invoked by the adversary and prepares the received stack
-pointer by checking that it has permission $\rwlx$.  The macro
-\texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} jumps to the
-capability in register $r$ in the manner described above. That is, it pushes
-local state (the contents of registers $\overline{r_{\var{priv}}}$) and the
-activation record (return code, return pointer, stack pointer) to the stack,
-creates an $\entry$ return pointer, restricts the stack pointer, clears the
-unused part of the stack, clears the necessary registers and jumps to $r$.  Upon
-return, the private state is restored. The macro \texttt{\footnotesize{mclear
-    $r$}} clears all the memory the capability in register $r$ has authority
-over.  The macro \texttt{\footnotesize{rclear $\var{regSet}$}} clears all the
-registers in $\var{regSet}$. The macro \texttt{\footnotesize{reqglob $r$}}
-checks whether the word in register $r$ is a global capability. The macro
+We define a number of reusable macros capturing the calling convention as well
+as other conveniences. The macros that utilise the stack assume that it is
+available in register $r_\stk$.
+
+The macro \texttt{\footnotesize{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} captures the
+parts of the calling convention related to actually transferring control to some
+adversarial code. Specifically, it pushes the contents of the private registers,
+$\overline{r_{\var{priv}}}$, to the stack. Then it pushes the restoration code
+to the stack. The restoration code will be executed as the first thing upon
+return, where it restores the stack pointer and the old program pointer. After
+the restoration code is pushed to the stack, a copy of the pc is adjusted to
+point to the first instruction after the control has been transferred and then
+pushed to the stack. From the stack pointer, a protected return pointer is
+created by adjusting it to point to the return code and encapsulating it by
+restricting its permission to enter. Next, the stack pointer is restricted to
+the unused part, and the unused part of the stack is cleared, so it does not
+govern any part of the private stack frame which means that it can safely be
+given away. Finally, the non-argument registers are cleared and we jump to $r$.
+Upon return after the restoration code has been executed, the restoration code
+is popped from the stack and the private words we pushed to the stack before the
+call are popped to the private registers. The implementation of \texttt{scall}
+is presented in \figurename~\ref{fig:scall-impl}, and the restoration code is
+presented in \figurename~\ref{fig:scall-rest-code}. The implementation of
+\texttt{scall} uses some of the macros we present later in this section.
+
+The macro \texttt{\footnotesize{mclear $r$}} clears all the memory the
+capability in register $r$ has authority over. It is used by \texttt{scall} to
+clear the unused part of the stack before control is transferred. It should also
+be used to clear the stack before returning to adversarial code. Similarly, the
+macro \texttt{\footnotesize{rclear $R$}} clears all the registers in the set
+$R$. It is also used by \texttt{scall}, and it should also be used before
+returning to adversarial code.
+
+The macros \texttt{\footnotesize{prepstack $r$}} and \texttt{\footnotesize{reqglob
+    $r$}} are the last macros related to the calling convention. The former,
+\texttt{prepstack} should be used when one receives a stack from an unknown
+sure. The \texttt{prepstack} macro first ensures the stack has permission
+read/write-local/execute, and then it adjusts the stack capability, so it
+follows the convention for the stack\footnote{The stack capability should always
+point to the top most word on the stack. A stack received from an unknown source
+can be treated as empty, so the stack capability should point just outside its
+range of authority.}. The other macro, \texttt{reqglob}, ensures that the
+capability in register $r$ is global. This macro should be used on callbacks
+received from unknown sources to ensure that they cannot be derived from the
+stack pointer.
+
+The remaining macros are not directly related to the calling convention, but
+they help making the examples in \sectionname~\ref{sec:examples} intelligible.
+The macros \texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}}
+add and remove elements from the stack. The macro \texttt{\footnotesize{fetch
+    $r$ $\var{name}$}} fetches the capability related to $\var{name}$ from the
+linking table and stores it in register $r$. The macro
 \texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}} allocates a closure
@@
 
-The technical appendix~\citep{technical_appendix} contains detailed descriptions
-of all the macros.
-\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
+The Appendix contains the implementation of all the macros used in
+\texttt{scall}. The technical appendix~\citep{technical_appendix} contains detailed descriptions
+of all the macros as well as all implementations.
+%\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
 
@@
   push $\encode(i_1)$
-  ...
+  push $\encode(i_2)$
+  push $\encode(i_3)$
   push $\encode(i_4)$
@@
 lemma statements. The examples may look deceivingly short, but it is because
-they use the macro instructions described in \sectionname~\ref{para:macros}.
+they use the macro instructions described in \sectionname~\ref{subsec:macros}.
 The examples would be
@@
 pointer\footnote{These assumptions are kept intentionally vague for brevity.
-  Full statements are in the technical appendix~\citep{technical_appendix}.}.
+  Full statements are in the Appendix.}.
 Furthermore

### 2018-05-02 fully-abstract-comp/ftlr-proof-cases.tex

@@
 \begin{lemma}[stackReadCondition works]
-  \label{lem:readcond-writecond-work}
+  \label{lem:stackreadcond-stackwritecond-work}
   If
@@
   \begin{itemize}
-  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
+  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
     From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
-    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:non-expansive}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
 
-  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
+  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
     The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by non-expansiveness of memory satisfaction, i.e.
@@
   \item $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$
-  \item $\perm \in \readAllowed{}$
-  \item $\npair{(c,c')}\in\lrv(H_\sigma,W)$
+  \item $\perm \in \readAllowed{}$ or $\perm' \in \readAllowed{}$
+  \item $\npair{(c,c')}\in\lrvg{\trust}(H_\sigma,W)$
   \item $W \oplus W_M$ is defined
   \item $n' < n$
+  \item ($\aaddr \in [\baddr,\eaddr]$ or $\aaddr' \in [\baddr',\eaddr']$)
   \item $w_S = \linCons{\ms_S(\aaddr)}$, $w_T = \linCons{\ms_T(\aaddr')}$
@@
   \item $W_M = W' \oplus W_M'$
-  \item $\memSat{(\ms_S,\stk,\ms_\stk\update{\aaddr}{w_S},\ms_T\update{\aaddr'}{w_T})}{W_M'}$
-  \item $\npair{(\ms_\stk(a),\ms_T(a))} \in \lrv(W')$
+  \item $\memSat[n']{(\ms_S,\stk,\ms_\stk\update{\aaddr}{w_S},\ms_T\update{\aaddr'}{w_T})}{W_M'}$
+  \item $\npair[n']{(\ms_\stk(a),\ms_T(a))} \in \lrv(W')$
   \end{itemize}
-  \dominique{11-3-2018: this lemma depends on Lemma~\ref{lem:non-linear-pure} for the non-linear case.}
 \end{lemma}
 \begin{proof}
-  From $\npair{(c,c')}\in\lrv(H_\sigma,W)$ with $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$, $\perm \in \readAllowed{}$ and $\perm' \in \readAllowed{}$, we get that $\perm = \perm'$, $\lin' = \linear$, $\baddr = \baddr'$, $\eaddr = \eaddr'$ and $\aaddr = \aaddr'$ and $\npair{(\baddr,\eaddr)} \in \stackReadCond{W}$.
+  From $\npair{(c,c')}\in\lrv(H_\sigma,W)$ with $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$, ($\perm \in \readAllowed{}$ or $\perm' \in \readAllowed{}$), we get that $\perm = \perm'$, $\lin' = \linear$, $\baddr = \baddr'$, $\eaddr = \eaddr'$ and $\aaddr = \aaddr'$ and $\npair{(\baddr,\eaddr)} \in \stackReadCond{W}$.
 
-  TODO: fix rest of proof
+  From Lemma~\ref{lem:stackreadcond-stackwritecond-work}, we then get that $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
 
-  Lemma~\ref{lem:readcond-writecond-work} then gives us a $W'$ and $W_M'$ such that $W_M = W' \oplus W_M'$ and $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$.
+  It remains to prove that $\memSat{(\ms_S,\stk,\ms_\stk\update{\aaddr}{w_S},\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+  We have to distinguish the case that $\isLinear{\ms_\stk(\aaddr)}$ and the opposite case.
+  \begin{itemize}
+  \item case $\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = 0$ and it follows from $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
+    From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm = \perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrvg{\trust}(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \stackWriteCond{W}$.
+    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:non-expansive}, we then get that $\memSat[n']{(\ms_S,\stk,\ms_\stk\update{\aaddr}{0},\ms_T\update{\aaddr'}{0})}{W_M'}$.
 
-  It remains to prove that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
-  We have to distinguish the case that $\isLinear{\ms_S(\aaddr)}$ and the opposite case.
-  \begin{itemize}
-  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
-    From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
-    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
-
-  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair{(c,c')}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
+  \item case $\neg\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = \ms_\stk(\aaddr)$ and it follows from $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
     The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by non-expansiveness of memory satisfaction, i.e.
@@
   \end{itemize}
-  TODO
 \end{proof}

### 2018-05-02 toplas/paper.tex
typos and minor changes
Added the malloc macro to the macro section.

@@
 \begin{toplas}
-  \begin{proof}
+  \begin{proof}[Proof sketch]
     Induction over $n$. By definition of $\stder(W)$, show
     $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
-    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg}$, and
+    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
     $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
@@
 linking table and stores it in register $r$. The macro
-\texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}} allocates a closure
-where $r$ points to the closure's code and a new environment is allocated (using
-malloc) where the contents of $\overline{r_i}$ is stored. In the code referred
-to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
+\texttt{\footnotesize{malloc $r$ $n$}} invokes malloc with size argument $n$.
+The \texttt{malloc} macro assumes that a capability for malloc resides in the
+linking table and is basically a fetch of the malloc capability followed by a
+setup of a return pointer. Finally, the macro \texttt{\footnotesize{crtcls
+    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
+closure's code and a new environment is allocated (using malloc) where the
+contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
+implicit fetch happens when an instruction refers to $x_i$.
 
@@
 \begin{figure}[htb]
+  \begin{toplas}
     \centering
@@
     \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
-    r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The restoration
-  code is given in \figurename~\ref{fig:scall-rest-code}. The variable
-  $\var{ofc}$ is the offset to the label after, and $\var{ofc_\var{act}} =
-  -5$ which is the offset to the first instruction of the activation record. The
-set $R = \RegName \setminus
+      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
+      restoration code is given in \figurename~\ref{fig:scall-rest-code}. The
+      variable $\var{ofc}$ is the offset to the label after, and
+      $\var{ofc_\var{act}} = -5$ which is the offset to the first instruction of
+      the activation record. The set $R = \RegName \setminus
       \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
     \label{fig:scall-impl}
+  \end{toplas}
+
 \end{figure}
 \begin{figure}[htb]
+  \begin{toplas}
     \centering
@@
 \end{lstlisting}
-  \caption{The restoration code used in \texttt{scall}. The variable $\var{ofc}$ is 5
-    which is the offset to the address where the old stack pointer is stored on
-    the stack.} 
+    \caption{The restoration code used in \texttt{scall}. The variable
+      $\var{ofc}$ is 5 which is the offset to the address where the old stack
+      pointer is stored on the stack.}
     \label{fig:scall-rest-code}
+  \end{toplas}
 \end{figure}

### 2018-05-03 toplas/paper.tex
Suggestion for macro section. Should we say something about temporary registers?
Added later definition for cofes to toplas appendix
Added text about malloc implementation

@@
 % Uncomment me to disable comments
-\renewcommand\lau[1]{}
-\renewcommand\dominique[1]{}
-\renewcommand\lars[1]{}
+%\renewcommand\lau[1]{}
+%\renewcommand\dominique[1]{}
+%\renewcommand\lars[1]{}
 
@@
 
-\itoplassug{Add the assumption that the malloc capability in examples
-  satisfies specification.}
-
 \subsection{Reusable macro instructions}
 \label{subsec:macros}
+The focus of this work is not to give a definitive specification and
+implementation of malloc which is why we do give a malloc implementation. We
+are, however, confident that it is possible to make an implementation of malloc
+that satisfies the malloc specification in Definition~\ref{spec:malloc}. There
+are in particular two simplifications in our system that makes things easy:
+First, we do not consider deallocation of memory which means that the data
+structure malloc uses to keep track of free memory does not have to handle
+reclaimed memory. Second, the malloc specification does nor permit malloc to run
+out of memory and thus prevent allocation. This is possible on our simple
+capability machine because it has an infinite address space. The capability with
+an infinite range of authority would of course need to be part of malloc, but it
+could also double as the data structure that keeps track of free memory.
+\itoplassug{If we have references for malloc being difficult, we could add that
+  to the text.}
+%\itoplassug{Add the assumption that the malloc capability in examples
+%  satisfies specification.}
+
 %\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
@@
 available in register $r_\stk$.
+\itoplassug{Lau, suggestion: should we mention temporary registers? They are
+  used in the macros, but it seems like a minor technical detail.}
 
@@
   where all Cauchy sequences have a limit.  
+\end{definition}
+
+\begin{definition}[Later on c.o.f.e.'s]
+  Let $\left( X, \left( \nequal \right)_{n=0}^{\infty}\right)$ be a \cofe{}. Define
+  \[
+  \blater \left( X, \left( \nequal \right)_{n=0}^{\infty}\right) = \left( X, \left(\nequal' \right)_{n=0}^{\infty}\right)
+  \] 
+  where for all $x$ and $x'$ in $X$
+  \[
+  x \nequal' x'\textit{  iff  } \begin{cases}
+    \top              & \text{if } n=0 \\
+    x \nequal[n-1] x' & \text{if } n > 0
+  \end{cases}
+  \]
+%  The set in the \cofe{} remains the same, but the equality relation is ``moved'' so it is equal at an earlier time.
+\todo[inline]{can we here say that to look under a later we have to use a computation step?}
 \end{definition}

### 2018-05-03 toplas/paper.tex
Updated FTLR proof, malloc, and macro section with Lars' suggestions

@@
 further in the next section.
-\itoplas{The specification of malloc is presented in \sectionname~\ref{sec:malloc-and-macros} as it uses the semantic model we build in Section~\ref{sec:logical-relation}.
+\itoplas{The specification of malloc is presented in \sectionname~\ref{sec:malloc} as it uses the semantic model we build in Section~\ref{sec:logical-relation}.
 For full details on the linking table, we refer to the technical
@@
     $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
-    $\mem'$, and $i \leq n'$ be given and assume $\Phi \step[i]
+    $\mem'$, and $i \leq n'$ be given and for
+    $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms \uplus
+    \ms_f)$ assume $\Phi \step[i]
     (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
@@
     configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
-    \var{conf}' \step[i-1] (\halted,\mem')$.
-
-    The rest of the proof roughly consists of two parts (1) argue that
-    $\var{conf}'$ is consistent with the world in the sense that the
-    register-file and memory still respect the world. (2) Argue that the rest of
-    the execution respects the world. The second case, relies on where the
-    $\pcreg$ in $\var{conf}'$ comes from. If the step to $\var{conf}'$ was a
-    jump, then use the fact that the register-file contains safe values in the
-    initial configuration. On the other hand, if the $\pcreg$ was just
-    incremented, then we need to apply the induction hypothesis which is
-    possible as all of the permission based conditions are downwards closed.
-    For instance in the case where the $\pcreg$ is read-execute, i.e.\ $\perm =
-    \exec$, we also assume $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
-    which entails $\npair[n'-1]{(\start,\addrend)} \in \readCond{}(\gl)(W)$ and
-    allows us to use the induction hypothesis.
-    In order to argue (1), we just need to consider the configurations the
-    operational semantics allows us to get to from the initial state. If we
-    consider the memory in $\var{conf}'$, then it either (a) remains unchanged
-    or (b) one address has been updated and in that case, the register-file contains an 
-    appropriate capability for writing. The latter case, (b) corresponds to the first
-    step executing a $\texttt{store}$ instruction, and the former (a) corresponds to
-    any other instruction.
-    In case (a), we can conclude that the memory still respects the world just
-    by downwards closure of memory satisfaction. In the other case, (b), we use
-    an auxiliary lemma called \emph{Conditions for \texttt{store} are
-      sufficient} (see the appendix, Lemma~\ref{lem:conds-store-stuff})
-    that states that given all the assumptions we have done so far, the updated
-    memory still satisfies the world.
-    All the possible changes to the register-file correspond to the effect of
-    some instruction, so for the non-trivial instructions we use a sufficiency
-    lemma.
+    \Phi' \step[i-1] (\halted,\mem')$.
+
+    The rest of the proof roughly consists of two parts: (1) Argue that $\Phi'$
+    is consistent with the world in the sense that the register-file and memory
+    still respect the world. (2) Argue that the rest of the execution respects
+    the world. The second case relies on where the $\pcreg$ in $\Phi'$ comes
+    from. If the step to $\Phi'$ was a jump, then use the fact that the
+    register-file contains safe values in $\Phi$. On the other hand, if the
+    $\pcreg$ was just incremented, then we need to apply the induction
+    hypothesis, which is possible as all of the permission based conditions are
+    downwards closed. For instance, in the case where the $\pcreg$ is
+    read-execute, i.e.\ $\perm = \exec$, we also assume
+    $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$ which entails
+    $\npair[n'-1]{(\start,\addrend)} \in \readCond{}(\gl)(W)$ and hence allows us to
+    use the induction hypothesis. In order to argue (1), we just need to
+    consider the configurations the operational semantics allows us to get to
+    from the initial state. If we consider the memory in $\Phi'$, then it either
+    (a) remains unchanged or (b) one address has been updated and in that case,
+    the register-file contains an appropriate capability for writing. The latter
+    case (b) corresponds to the first step executing a $\texttt{store}$
+    instruction, and the former (a) corresponds to any other instruction. In
+    case (a), we can conclude that the memory still respects the world just by
+    downwards closure of memory satisfaction. In the other case, (b), we use an
+    auxiliary lemma called \emph{Conditions for \texttt{store} are sufficient}
+    (see the appendix, Lemma~\ref{lem:conds-store-stuff}) that states that given
+    all the assumptions we have so far, the updated memory still satisfies
+    the world. All the possible changes to the register-file correspond to the
+    effect of some instruction, so for each non-trivial instruction we use a
+    lemma appropriate for that instruction.
 
@@
 \begin{toplas}
-  \section{Malloc and macros}
-  \label{sec:malloc-and-macros}
-With the semantic model defined, we would like to show its usefulness by proving
-the correctness of a series of examples that rely
-on well-bracketedness and local-state encapsulation and use the calling
-convention to enforce these properties. However, the programs that
-run on our capability are assembly programs and even program examples that would
-be small in a high-level language, they become big and unintelligible at this
-low level. In order to make our program examples intelligible, we introduce a
-series of low-level abstractions in the form of macros.
-
-To make the program examples realistic, they have access to a trusted malloc, so
-they can allocate new memory. Our results are not specific to
-any one implementation of malloc, so rather than providing the implementation,
-we provide a specification that malloc must satisfy.
-
-\subsection{Malloc specification}
-  \itoplassug{Suggestion: add the malloc specification}
-Before we give the malloc specification, we provide some intuition about its requirements.
+  \section{Malloc}
+  \label{sec:malloc}
+In the examples we present in \sectionname~\ref{sec:examples}, we will assume
+the existence of a trusted malloc routine so that it is possible to allocate new
+memory. Our semantic model is not specific to a particular implementation of
+malloc, so rather than providing the implementation, we provide a specification
+that malloc must satisfy. This section can be skipped on first read through.
 
@@
 piece of memory not just under the control of malloc. We also need to make sure
-that malloc actually behave as intended. That is when malloc is invoked with a
+that malloc actually behaves as intended, that is, when malloc is invoked with a
 length argument, then it actually returns a capability for a fresh piece of
-memory of that size. It should be fresh in the sense that malloc hasn't given
-out a capability for any part of that memory before and won't do it in the
+memory of that size. It should be fresh in the sense that malloc has not given
+out a capability for any part of that memory before and will not do it in the
 future. Finally, we also need to make sure that malloc behaves correctly when
@@
 
-In order to express the semantic requirements on malloc, the specification will
+For the semantic requirements on malloc, the specification will
 require the existence of some initial region for malloc. Malloc does not
@@
 contains the malloc region. In other words, the remaining regions in the world
-do not influence which regions malloc accepts. Malloc should retain this
+should not influence which regions malloc accepts. Malloc should retain this
 property throughout execution, so it must hold true for any private future
@@
 Malloc can be invoked with either a valid or an invalid length argument. In the
-case where malloc is invoked with an invalid argument, that is a negative
-integer or a capability, we require malloc to simply fail. This part of the
-specification does not rely on the malloc region as failing is considered
+case where malloc is invoked with an invalid argument, that is, a negative
+integer or a capability, we simply require malloc to fail. This part of the
+specification does not rely on the malloc region, as failing is considered
 safe behaviour. When malloc is invoked with a valid length, then malloc should
 return a fresh memory segment of the correct length.
-\todo[inline]{Everything below this is work in progress..}
 Malloc may have been invoked a number of times before any given invocation, so
-the internal data structures of malloc may be in a different state than it
-started in.
+the internal data structures of malloc may be in a different state than they
+were initially.
 We know that the memory of malloc is in a state consistent with the
@@
 after the call. This allows the caller to keep private capabilities in the
-register file. This is strictly speaking not necessary as we can use our calling
-convention to protect our local state.
+register file (strictly speaking this is not necessary as we can use our calling
+convention to protect our local state).
 
@@
 \end{definition}
-As described previously, the specification of malloc ensure that malloc has no
+As described previously, the specification of malloc ensures that malloc has no
 capabilities pointing \emph{out} of malloc. It does, however, not say anything about
-capabilities that point \emph{in} to malloc. We obviously can obviously not
+capabilities that point \emph{in} to malloc. We obviously can not
 allow this if we want to trust malloc. We have chosen to keep the malloc
@@
   \label{lem:malloc-in-vr}
-  For $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
+  For all $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
@@
 
-\subsection{Reusable macro instructions}
-\label{subsec:macros}
 The focus of this work is not to give a definitive specification and
@@
 
+\section{Reusable macro instructions}
+\label{sec:macros}
 %\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
@@
 I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
+With the semantic model defined, we would like to show its usefulness by proving
+the correctness of a series of examples that rely
+on well-bracketedness and local-state encapsulation and use the calling
+convention to enforce these properties. However, the programs that
+run on our capability are assembly programs and even program examples that would
+be small in a high-level language become big and unintelligible at this
+low level. Thus to make our program examples intelligible, we introduce a
+series of low-level abstractions in the form of macros.
 We define a number of reusable macros capturing the calling convention as well
@@
 The macro \texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} captures the
-parts of the calling convention related to actually transferring control to some
-adversarial code. Specifically, it pushes the contents of the private registers,
-$\overline{r_{\var{priv}}}$, to the stack. Then it pushes the restoration code
-to the stack. The restoration code will be executed as the first thing upon
-return, where it restores the stack pointer and the old program pointer. After
-the restoration code is pushed to the stack, a copy of the pc is adjusted to
-point to the first instruction after the control has been transferred and then
-pushed to the stack. From the stack pointer, a protected return pointer is
-created by adjusting it to point to the return code and encapsulating it by
-restricting its permission to enter. Next, the stack pointer is restricted to
-the unused part, and the unused part of the stack is cleared, so it does not
-govern any part of the private stack frame which means that it can safely be
-given away. Finally, the non-argument registers are cleared and we jump to $r$.
-Upon return after the restoration code has been executed, the restoration code
-is popped from the stack and the private words we pushed to the stack before the
-call are popped to the private registers. The implementation of \texttt{scall}
-is presented in \figurename~\ref{fig:scall-impl}, and the restoration code is
-presented in \figurename~\ref{fig:scall-rest-code}. The implementation of
-\texttt{scall} uses some of the macros we present later in this section.
-
-The macro \texttt{\footnotesize{mclear $r$}} clears all the memory the
-capability in register $r$ has authority over. It is used by \texttt{scall} to
-clear the unused part of the stack before control is transferred. It should also
-be used to clear the stack before returning to adversarial code. Similarly, the
-macro \texttt{\footnotesize{rclear $R$}} clears all the registers in the set
-$R$. It is also used by \texttt{scall}, and it should also be used before
-returning to adversarial code.
-
-The macros \texttt{\footnotesize{prepstack $r$}} and \texttt{\footnotesize{reqglob
-    $r$}} are the last macros related to the calling convention. The former,
-\texttt{prepstack} should be used when one receives a stack from an unknown
-sure. The \texttt{prepstack} macro first ensures the stack has permission
-read/write-local/execute, and then it adjusts the stack capability, so it
-follows the convention for the stack\footnote{The stack capability should always
-point to the top most word on the stack. A stack received from an unknown source
-can be treated as empty, so the stack capability should point just outside its
-range of authority.}. The other macro, \texttt{reqglob}, ensures that the
-capability in register $r$ is global. This macro should be used on callbacks
-received from unknown sources to ensure that they cannot be derived from the
-stack pointer.
-
-The remaining macros are not directly related to the calling convention, but
-they help making the examples in \sectionname~\ref{sec:examples} intelligible.
-The macros \texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}}
-add and remove elements from the stack. The macro \texttt{\footnotesize{fetch
-    $r$ $\var{name}$}} fetches the capability related to $\var{name}$ from the
-linking table and stores it in register $r$. The macro
-\texttt{\footnotesize{malloc $r$ $n$}} invokes malloc with size argument $n$.
-The \texttt{malloc} macro assumes that a capability for malloc resides in the
-linking table and is basically a fetch of the malloc capability followed by a
-setup of a return pointer. Finally, the macro \texttt{\footnotesize{crtcls
-    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
-closure's code and a new environment is allocated (using malloc) where the
-contents of $\overline{r_i}$ is stored. In the code referred to by $r$, an
-implicit fetch happens when an instruction refers to $x_i$.
-
-The Appendix contains the implementation of all the macros used in
-\texttt{scall}. The technical appendix~\citep{technical_appendix} contains detailed descriptions
-of all the macros as well as all implementations.
-%\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} captures
+those parts of the calling convention related to actually transferring control
+to some adversarial code. Specifically, it pushes the contents of the private
+registers, $\overline{r_{\var{priv}}}$, to the stack. Then it pushes the
+``restoration'' code to the stack. The restoration code will be executed as the
+first thing upon return, and restores the stack pointer and the old program
+pointer. After the restoration code is pushed to the stack, a copy of the pc is
+pushed to the stack after it has been adjusted to point to the first instruction
+after the jump. From the stack pointer, a protected return pointer is created by
+adjusting it to point to the return code and encapsulating it by restricting its
+permission to Enter. Next, the stack pointer is restricted to the unused part,
+and the unused part of the stack is cleared, so it does not govern any part of
+the private stack frame which means that it can safely be given away. Finally,
+the non-argument registers are cleared and we jump to $r$. Upon return after the
+restoration code has been executed, the restoration code is popped from the
+stack and the private words we pushed to the stack before the call are popped to
+the private registers. The implementation of \texttt{scall} is presented in
+\figurename~\ref{fig:scall-impl}, and the restoration code is presented in
+\figurename~\ref{fig:scall-rest-code}. The implementation of \texttt{scall} uses
+some of the macros we now present.
 
@@
 
+The macro \texttt{\footnotesize{mclear $r$}} clears all the memory the
+capability in register $r$ has authority over. It is used by \texttt{scall} to
+clear the unused part of the stack before control is transferred. It should also
+be used to clear the stack before returning to adversarial code. Similarly, the
+macro \texttt{\footnotesize{rclear $R$}} clears all the registers in the set
+$R$. It is also used by \texttt{scall}, and it should also be used before
+returning to adversarial code.
+
+The macros \texttt{\footnotesize{prepstack $r$}} and \texttt{\footnotesize{reqglob
+    $r$}} are the last macros related to the calling convention. The former,
+\texttt{prepstack} should be used when one receives a stack from an unknown
+source. The \texttt{prepstack} macro first ensures the stack has permission
+read/write-local/execute, and then it adjusts the stack capability, so it
+follows the convention for the stack\footnote{The stack capability should always
+point to the top most word on the stack. A stack received from an unknown source
+can be treated as empty, so the stack capability should point just outside its
+range of authority.}. The other macro, \texttt{reqglob}, ensures that the
+capability in register $r$ is global. This macro should be used on callbacks
+received from unknown sources to ensure that they cannot be derived from the
+stack pointer.
+
+The remaining macros are not directly related to the calling convention, but
+they help making the examples in \sectionname~\ref{sec:examples} intelligible.
+The macros \texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}}
+add and remove elements from the stack. The macro \texttt{\footnotesize{fetch
+    $r$ $\var{name}$}} fetches the capability related to $\var{name}$ from the
+linking table and stores it in register $r$. The macro
+\texttt{\footnotesize{malloc $r$ $n$}} invokes malloc with size argument $n$.
+The \texttt{malloc} macro assumes that a capability for malloc resides in the
+linking table and is basically a fetch of the malloc capability followed by a
+setup of a return pointer. Finally, the macro \texttt{\footnotesize{crtcls
+    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
+closure's code and a new environment is allocated (using malloc) and the
+contents of $\overline{r_i}$ is stored in the environment. In the code referred
+to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
+
+The Appendix contains the implementation of all the macros used in
+\texttt{scall}. The technical appendix~\citep{technical_appendix} contains more
+detailed descriptions of all the macros as well as all implementations.
+%\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
+We stress that the macros correspond to series of instructions as seen in
+\figurename~\ref{fig:scall-impl} and \ref{fig:scall-rest-code}, and that they
+are introduced for intelligibility. The examples of
+\sectionname~\ref{sec:examples}, the program examples are stated using the
+macros, but the proofs work on the expanded macros.
 \end{toplas}
@@
 lemma statements. The examples may look deceivingly short, but it is because
-they use the macro instructions described in \sectionname~\ref{subsec:macros}.
+they use the macro instructions described in \sectionname~\ref{sec:macros}.
 The examples would be

### 2018-05-04 toplas/paper.tex
Added first draft for section about reasoning

@@
 % Conv defs
+\newcommand{\linksto}[4]{\ensuremath{#1} \text{ links } \ensuremath{#2} \text{ as } \ensuremath{#3} \text{ to } \ensuremath{#4}}
 \newcommand{\lookingat}[3]{\ensuremath{#1} \text{ is looking at } \ensuremath{#2} \text{ followed by } \ensuremath{#3}}
@@
   move r_0 r_stk
-  lea r_0 $\var{ofc_\var{act}}$ // -5 is the offset to the first instruction of the activation code
+  lea r_0 $\var{ofc_\var{rec}}$ // -5 is the offset to the first instruction of
+the recovery code
   restrict r_0 $\encodePermPair((\local,\entry))$
@@
       variable $\var{ofc}$ is the offset to the label after, and
-      $\var{ofc_\var{act}} = -5$ which is the offset to the first instruction of
+      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
       the activation record. The set $R = \RegName \setminus
@@
 \begin{toplas}
-\section{Reasoning about programs}
+\section{Reasoning about programs on a capability machine}
 \label{sec:reasoning}
@@
 
+% Anti-reduction lemma
+Out capability machine only allows the final memory of an execution to be
+observed, so naturally the correctness lemmas we prove in
+\sectionname~\ref{sec:examples} are statements about the memory in the halted
+configuration. In order to prove something about the final memory, we encode it
+as a region in a world and show that the initial configuration is in the
+$\observations$-relation w.r.t.\ that world. The $\observations$-relation says
+that if the initial configuration halts successfully, then the memory in the
+halting state must still respect a private future world of the initial world. It
+is, however, bothersome and error prone to try to argue about the entire
+execution in one go. Instead we want to reason modularly, in the sense that we
+only want to reason about parts of the execution at a time. To this end, we
+prove an anti-reduction lemma that essentially says that if we can show that
+an initial configuration, $\Phi$, steps to a configuration $\Phi'$ and
+that $\Phi'$ is in the $\observations$-relation, then $\Phi$ is also in the
+observation relation.
+\begin{lemma}[Anti-reduction for $\observations$]
+\label{lem:anti-red-obs}
+  \begin{align*}
+    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
+    & \quad n' \geq n - i \land W' \futurestr W \land \\
+    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
+    & \quad \npair[n']{(\reg',\ms')} \in \observations(W') \\
+    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
+  \end{align*}
+\end{lemma}
+
+% FTLR
+The anti-reduction works well when we reason about the execution of known code,
+because we can execute it. When we want to reason about unknown code, the
+anti-reduction lemma does not apply because we do not know what instructions are
+being executed. This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
+As a reminder, the FTLR says that if a capability only has access to safe values
+with respect to a world, then it is safe to use it for execution in the same
+world. The unknown code we consider will be assumed to only have access to safe
+values, which typically means that we assume it has access to a linking table
+with a malloc capability and otherwise consists of instructions. This allows us
+to use the FTLR seeing as Malloc is a safe value cf.
+Lemma~\ref{lem:malloc-in-vr} and instructions are really just integers and they
+are always safe values. Here, it is important to remember that the FTLR gives us
+that the capability for the untrusted code is safe w.r.t. some world, so even
+though we described the usual assumption on the unknown code's memory, it is the
+semantic model of the memory, i.e.\ the world, that the FTLR considers. In order
+to argue that a specific configuration is safe, we need to argue that the
+configuration is safe with respect to the world. That is, we need to argue that
+the memory satisfies the world and the register-file is in the
+$\stdrr$-relation. For instance, in the case where the untrusted code assumes
+control first, we will use the FTLR to show that the capability for the unknown
+code can be used for execution. The unknown code will get access to some known,
+trusted code through a capability in the initial register-file. Because the
+capability is in the register-file, we will have to argue that it is safe which
+we do using Lemma~\ref{lem:anti-red-obs}.
+
+% Many details to get right when reasoning about low-level programs.
+There are many details to get right when programming in assembly. These details
+carry over to proofs about assembly programs, so the proofs become cumbersome.
+It is especially annoying, when the same line of reasoning is applied in
+multiple places. For instance, in $\texttt{scall}$ the setup before the jump to
+the call is always the same and the recovery upon return is also always the
+same, and in both cases the reasoning will be the same every time.
+We deal with this part of the reasoning in the following lemma.
+\begin{lemma}[$\mathtt{scall}$ works]
+  \label{lem:scall-works}
+  If
+  \begin{enumerate}
+  \item \label{item:scall:memsat} $\memSat[n]{\ms}{\revokeTemp{W}}$ 
+  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
+  \item \label{item:scall:stack} $(\reg,\ms) \text{ is looking at }
+    \mathtt{scall}\;r(\overline{r_{\mathit{arg}}},
+    \overline{r_{\mathit{priv}}}) \text{ followed by }
+    c_{\mathit{next}}$\footnote{Defined in the appendix.}
+  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
+  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} If
+    \begin{itemize}
+    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{rec}}
+      \uplus \ms_{\mathit{unused}}')$,
+    \item $W' =
+      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
+    \item $\memSat[n-1]{\ms''}{W'}$
+    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
+    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
+      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
+      \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r \mapsto \reg(r)]$ 
+    \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
+    \item $\npair[n-1]{c_\stk} \in \stdvr(W')$
+    \end{itemize}
+
+    then we have that $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
+  \item \label{item:scall:hyp-cont} \emph{Hyp-Cont} If
+    \begin{itemize}
+    \item $n' \leq n-2$
+    \item $W'' \futurewk \revokeTemp{W}$
+    \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
+    \item for all $r$, we have that:
+      \begin{equation*}
+        \reg'(r)
+        \begin{cases}
+          = c_{\mathit{next}} &\text{ if } r = \pcreg\\
+          = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
+          \in \stdvr(\revokeTemp{W''}) &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\}
+        \end{cases}
+      \end{equation*}
+    \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
+    \end{itemize}
+
+    then we have that $\npair[n']{(\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')} \in \observations(W'')$
+  \end{enumerate}
+  Then 
+  \begin{itemize}
+    \item $\npair{(\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}})} \in \observations(W)$
+  \end{itemize}
+
+  % Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
+  % words for stack pointer and return pointer and some fixed number of words for
+  % the standard return instructions).
+\end{lemma}
+The essence of the \texttt{scall} lemma is to show \texttt{scall} is executed in
+a reasonable state (\ref{item:scall:memsat}-\ref{item:scall:stack}), it is safe
+to actually transfer control to the callee (\ref{item:scall:hyp-callee}), and that
+the return capability given to the callee is safe (\ref{item:scall:hyp-cont}),
+which corresponds to showing that the remainder of the program is safe to execute.
+
+In the common case \texttt{scall} is used to invoke untrusted code, and we will
+only have very general assumptions about that code such as it has contains a
+capability for a linking table with malloc in it and all other addresses are
+instructions. These values are all safe which allows us to conclude that the
+read condition holds for the untrusted code which allows us to use the
+Theorem~\ref{thm:ftlr} to conclude that it can be safely executed if the memory
+and the register-file are safe. By assumption in Hyp-Callee, the memory is safe,
+so it suffices to show that the register-file contains safe values, which
+amounts to showing that the arguments in the call are safe.
+
+By using Lemma~\ref{lem:scall-works} to reason about \texttt{scall}, the proofs
+become agnostic to the actual implementation of \texttt{scall}. In other words,
+should we change the implementation of \texttt{scall}, then we just need to
+prove that Lemma~\ref{lem:scall-works} holds for the new implementation in order
+to gain that all our results still hold true.
+
+The \texttt{malloc} instruction is also used often, so we also prove
+Lemma~\ref{lem:malloc-works} to help reason about it. The structure of the
+\texttt{malloc} lemma is close to that of the \texttt{scall} lemma. It also has
+requirements on the configuration just before \texttt{malloc} is executed
+(\ref{item:malloc:looking-at}-\ref{item:malloc:malloc-sat}) as well as
+requirements on the execution afterwards (\ref{item:malloc:hyp-cont}). It does,
+however, not have any requirements on the callee as we can rely on the
+\texttt{malloc} specification to reason about this part once and for all.
+\begin{lemma}[$\mathtt{malloc}$ works]
+  \label{lem:malloc-works}
+  If
+  \begin{enumerate}
+  \item \label{item:malloc:looking-at}\lookingat{(\reg,\ms)}{\mathtt{malloc}\;r\;k}{c_{\var{next}}}
+  \item $k \geq 0$
+  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
+  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
+  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
+  \item $\memSat{\ms}{W}$
+  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
+  \item \label{item:malloc:malloc-sat} $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
+  \item \label{item:malloc:hyp-cont} \emph{Hyp-Cont}
+    If
+    \begin{itemize}
+    \item $n' \leq n-1$
+    \item $\iota_\malloc \futurewk W(i)$
+    \item $\memSat[n']{\ms_{\var{footprint}}' \uplus \ms'}{W[i \mapsto \iota_\malloc]}$
+    \item $\memSat[n']{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
+      \begin{equation*}
+        \reg'(r') = 
+        \begin{cases}
+          c_{\var{next}} & r' = \pcreg \\
+          ((\rwx,\glob),\start,\addrend,\addr) & r' = r \\
+          \reg(r) & r' \not\in \RegName_t \union \{\pcreg, r, r_1\}
+        \end{cases}
+      \end{equation*}
+    \item $\addrend - \start = k - 1$
+    \item $\dom(\ms_{\var{alloc}}) = [\start,\addrend]$
+    \item $\forall \addr \in [\start,\addrend]\ldotp \ms_{\var{alloc}}(\addr) = 0$
+    \end{itemize}
+    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}}'  \uplus \ms_{\var{alloc}})} \in \observations(W[\iota_\malloc])$
+  \end{itemize}
+  Then
+  \[
+    \npair{(\reg,\ms)} \in \observations(W)
+  \]
+\end{lemma}
+In the technical appendix~\citep{technical_appendix}, we also define a lemma for
+the macro used to create closures, \texttt{crtcls}. We stress that these lemmas
+are conveniences and are not strictly necessary to reason about programs for the
+capability machine. In principle, one could reason solely in the semantic model
+without the use of these lemmas, but that would lead to a lot of repetition.
 \end{toplas}
@@
 
+\subsection{Reasoning about programs definitions}
+\begin{definition}
+  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ 
+  iff
+  \begin{itemize}
+  \item $\reg(\pcreg) = ((p,g),b,e,a)$
+  \item $p = \rwx$, $p = \exec$, or $p = \rwlx$
+  \item $a+n\leq e$, $b\leq a\leq e$
+  \item $\ms(a+0,\cdots,a+n) = [i_0,\cdots,i_n]$
+  \item $c_{\mathit{next}} = ((p,g),b,e,a+n+1)$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}
+  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c_\malloc}'' 
+  iff
+  \begin{itemize}
+  \item $\reg(pc) = \stdcap$
+  \item $\ms(\start) = ((\_,\_),\start_\link,\_,\_)$
+  \item $\ms(\start_\link+j) = c$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}
+  We say that $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
+  iff
+  \begin{itemize}
+  \item $\reg(r_\stk) =((\rwlx,\local),b_\stk,e_\stk,a_\stk)$
+  \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
+  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$ \lau{Maybe make it clear what happens when $\ms_\stk$ is empty}
+  \item $b_\stk - 1\leq a_\stk$
+  \end{itemize}
+\end{definition}
+
 \end{document}

### 2018-05-05 toplas/paper.tex
Fixed compilation
Some minor rephrasings, restructure the FTLR proof sketch a bit.

@@
 
+\dominique{Follow order of figure in discussion?}
 The $\load{}{}$ instruction loads the contents of the requested memory location
@@
 in software, precisely to allow experimenting with alternative models like ours.
+\dominique{Update to CHERI's new ccall story.}
 
@@
   linking model)\\}
-\itoplassug{Suggestion: We also have the flag table ussed
+\itoplassug{Suggestion: We also have the flag table used
   for assertions, maybe we should add a description of it here.\\}
@@
   the callee could be well-behaved and wish to ensure well-bracketed calls and
-  local-state encapsulation. The calling convention puts no restriction on the
-  callee that the caller itself does not follow, so by following the calling
-  convention, the callee would also get well-bracketedness and local-state
-  encapsulation. Generally, the calling convention ensures the two properties
-  for anyone who employs it, so it scales to the setting where multiple
-  distrusting parties need to invoke each other and each of them can protect
-  themself by using the calling convention.
+  local-state encapsulation as well. The calling convention puts no restriction
+  on the callee that the caller itself does not follow, so by following the
+  calling convention, the callee can also obtain those guarantees. In other
+  words, the calling convention is modular and scales to scenario's with multiple
+  distrusting parties invoking each other.
 \end{toplas}
@@
 In contrast, enter capabilities are opaque and can only be used to jump to the
-address they point to.
-\begin{toplas}
-  As enter capabilities are the only capabilities that have to use the address they point to, it is natural that the $\entryCond{}$ is also the only permission-based condition that depends on the current address of the capability.
+address they point to.%
+\begin{toplas}%
+  This is why the $\entryCond{}$ depends on the current address of the capability, unlike for other types of capabilities.
 \end{toplas}
@@
     (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
-    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halting
+    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halted
     configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
@@
 
-    The rest of the proof roughly consists of two parts: (1) Argue that $\Phi'$
-    is consistent with the world in the sense that the register-file and memory
-    still respect the world. (2) Argue that the rest of the execution respects
-    the world. The second case relies on where the $\pcreg$ in $\Phi'$ comes
-    from. If the step to $\Phi'$ was a jump, then use the fact that the
-    register-file contains safe values in $\Phi$. On the other hand, if the
-    $\pcreg$ was just incremented, then we need to apply the induction
-    hypothesis, which is possible as all of the permission based conditions are
-    downwards closed. For instance, in the case where the $\pcreg$ is
-    read-execute, i.e.\ $\perm = \exec$, we also assume
-    $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$ which entails
-    $\npair[n'-1]{(\start,\addrend)} \in \readCond{}(\gl)(W)$ and hence allows us to
-    use the induction hypothesis. In order to argue (1), we just need to
-    consider the configurations the operational semantics allows us to get to
-    from the initial state. If we consider the memory in $\Phi'$, then it either
-    (a) remains unchanged or (b) one address has been updated and in that case,
-    the register-file contains an appropriate capability for writing. The latter
-    case (b) corresponds to the first step executing a $\texttt{store}$
-    instruction, and the former (a) corresponds to any other instruction. In
-    case (a), we can conclude that the memory still respects the world just by
-    downwards closure of memory satisfaction. In the other case, (b), we use an
-    auxiliary lemma called \emph{Conditions for \texttt{store} are sufficient}
-    (see the appendix, Lemma~\ref{lem:conds-store-stuff}) that states that given
-    all the assumptions we have so far, the updated memory still satisfies
-    the world. All the possible changes to the register-file correspond to the
-    effect of some instruction, so for each non-trivial instruction we use a
-    lemma appropriate for that instruction.
+    The rest of the proof considers the different ways the step $\Phi \step
+    \Phi'$ could have occurred, depending on the instruction being executed. For
+    each of these cases, we argue (1) that $\Phi'$ is consistent with the world
+    in the sense that the register-file and memory still respect the world and
+    (2) that the rest of the execution respects the world. Depending on where
+    the $\pcreg$ in $\Phi'$ comes from, the second result is proven in one of
+    two ways. If the step to $\Phi'$ was a jump, then the new $\pcreg$ is one of
+    the safe values in $\Phi$'s registers and the value relation can be used to
+    argue that those can be safely jumped to. On the other hand, if the $\pcreg$
+    was just incremented, then we can apply the induction hypothesis.
+
+    In order to argue (1), we consider what configurations the operational
+    semantics allows us to get to from the initial state. If we consider the
+    memory in $\Phi'$, then it either (a) remains unchanged or (b) one address
+    has been updated and in that case, the register-file contains an appropriate
+    capability for writing. The latter occurs when the instruction being
+    executed is a $\texttt{store}$, and otherwise the former. In case (a), we
+    can conclude that the memory still respects the world just by downwards
+    closure of memory satisfaction. In case (b), we use an auxiliary lemma that
+    uses the safety of the write capability used by the $\texttt{store}$
+    instruction to show that the updated memory satisfies the world. To show
+    that the updated register-file is safe, we consider the changes made to it
+    by all instructions in separate lemmas and show that they all preserve
+    safety of the register file.
 
@@
     Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}}'  \uplus \ms_{\var{alloc}})} \in \observations(W[\iota_\malloc])$
-  \end{itemize}
+  \end{enumerate}
   Then
@@
 
-\lau{Reviewer C, esop, would like a better comparison with \citec{Devriese:2016ObjCap} (but also admits that they did not follow the below discussion).}
+\lau{Reviewer C, esop, would like a better comparison with \citet{Devriese:2016ObjCap} (but also admits that they did not follow the below discussion).}
 Our logical relation is a unary, step-indexed Kripke logical relation with

### 2018-05-05 toplas/paper.tex
Some modifications to the malloc section, explain that we use the LR machinery in the spec to allow malloc to depend on evolving internal state, describe some things a bit less defensively.

@@
   In the examples we present in \sectionname~\ref{sec:examples}, we will assume
-the existence of a trusted malloc routine so that it is possible to allocate new
-memory. Our semantic model is not specific to a particular implementation of
-malloc, so rather than providing the implementation, we provide a specification
-that malloc must satisfy. This section can be skipped on first read through.
-
-Malloc is assumed to be part of the trusted computing base, as mentioned in
-\sectionname~\ref{sec:capab-mach-with}. A program with access to malloc should
-be able to retain this access even if it invokes code it does not trust. In
-order to achieve this, a capability for malloc must be global. It is also
-important that malloc remains trustworthy to all entities on the machine
-throughout execution. That is, malloc should encapsulate its local state and
-not depend on any external code. To make sure that malloc cannot be tampered
-with directly through the malloc capability, we require it to have enter
-permission.
-
-The requirements so far are syntactic and prevent direct tampering in the sense
-that the capability intended for invoking malloc only can be used for
-invocation. This does, however, not prevent the implementation of malloc from
-colluding with other code. Specifically, malloc could depend on the state of a
-piece of memory not just under the control of malloc. We also need to make sure
-that malloc actually behaves as intended, that is, when malloc is invoked with a
-length argument, then it actually returns a capability for a fresh piece of
-memory of that size. It should be fresh in the sense that malloc has not given
-out a capability for any part of that memory before and will not do it in the
-future. Finally, we also need to make sure that malloc behaves correctly when
-given a nonsensical length argument such as a negative integer or a capability.
-These requirements relate to the behaviour of malloc, so we cannot hope to
-express them syntactically. Instead we will use the semantic model defined in
-Section~\ref{sec:logical-relation} to express this part of the specification.
-
-For the semantic requirements on malloc, the specification will
-require the existence of some initial region for malloc. Malloc does not
-disappear from the system, so the malloc region must be permanent. This also
-means that the malloc specification should guarantee that use of malloc is safe
-at any point of execution no matter who has invoked it. To this end, the
-requirement on the initial malloc region should also hold true in any private
-future world.
+  the existence of a trusted malloc routine so that both the trusted code and
+  the adversary can allocate new memory. Malloc is considered part of the
+  trusted computing base, as mentioned in
+  \sectionname~\ref{sec:capab-mach-with}. This is unavoidable: if we cannot
+  trust malloc, then we cannot use the memory it allocates as we have no idea
+  whether it is aliased by some untrusted program.
+
+  Our semantic model is not specific to a particular implementation of malloc,
+  so rather than providing the implementation, we provide a specification that
+  malloc must satisfy. The specification is a bit technical, but really just
+  expresses standard expectations of how malloc should behave and this section
+  can safely be skipped on first read.
+
+We require a global capability for
+invoking malloc (because if the capability were local, then a program with
+access to malloc would have to give up this access when invoking untrusted
+code). The capability is assumed to have enter permission, so that malloc can
+protect its internal state even when the capability is shared with untrusted
+code.
+
+In addition to these syntactic requirements, we also specify standard intuitive
+expectations for how malloc behaves. When malloc is invoked with a length
+argument, then it will return a capability for a fresh piece of memory of that
+size. It should be fresh in the sense that malloc has not already given out a
+capability for any part of that memory before and will not do so in the future.
+Finally, we also need to make sure that malloc behaves correctly when given a
+nonsensical length argument such as a negative integer or a capability. However,
+formulating these requirements is harder than one might expect. The problem is
+that we want to allow malloc implementations to rely on internal state and
+invariants on that state and it is reasonable to expect both this internal state
+and the invariants to change over time. To express that malloc should behave
+according to its specifications, but only if its internal state satisfies its
+current invariants in an executing system, we will use the semantic model
+defined in Section~\ref{sec:logical-relation} to express our specification.
+
+To allow malloc implementations to rely on internal state and invariants for
+that state, we assume an initial region for malloc. The region is assumed to be
+permanent (since safety of the global malloc capability will depend on its
+presence).
+% DOMI: I don't understand what the following uncommented text means
+% This also means that the malloc specification should guarantee that
+% use of malloc is safe at any point of execution no matter who has invoked it. To
+% this end, the requirement on the initial malloc region should also hold true in
+% any private future world.
 % priv future means that malloc should not depend on local capabilities?
-
-The implementation of malloc should not depend on any memory outside the
-implementation of malloc. This means that malloc should not have any
-capabilities in its internals that govern other programs. As explained in
-\sectionname~\ref{subsec:worlds}, a region's state interpretation function
-determines what memory segments it permits with respect to some world. The state
-interpretation function of malloc should permit the same regions with respect to
-a world with the malloc region as it would with respect to a world that only
-contains the malloc region. In other words, the remaining regions in the world
-should not influence which regions malloc accepts. Malloc should retain this
-property throughout execution, so it must hold true for any private future
-region of the initial malloc region.
-
-The malloc specification will also dictate what malloc should do when invoked.
-Malloc can be invoked with either a valid or an invalid length argument. In the
-case where malloc is invoked with an invalid argument, that is, a negative
-integer or a capability, we simply require malloc to fail. This part of the
-specification does not rely on the malloc region, as failing is considered
-safe behaviour. When malloc is invoked with a valid length, then malloc should
-return a fresh memory segment of the correct length.
-Malloc may have been invoked a number of times before any given invocation, so
-the internal data structures of malloc may be in a different state than they
-were initially.
-We know that the memory of malloc is in a state consistent with the
-protocol dictated by the initial malloc region in the sense that we can assume
-that malloc is invoked in a configuration where part of it satisfies a private
-future region of the initial malloc region. Knowing the state of the malloc
-region allows us to express what we mean by a \emph{fresh} region. This means
-that malloc should not keep any capabilities for the freshly allocated region.
-To ensure this, we require the footprint of malloc after the call to be disjoint
-from the freshly allocated memory. Further, the footprint should satisfy a
-public future region of the malloc region the invocation happened in. This means
-that it cannot depend on the freshly allocated region. It further allows
-subsequent calls to malloc as it provides one of the requirements in the malloc
-specification.
-
-Malloc is part of the trusted computing base. If we cannot trust malloc, then we
-cannot use the memory it allocates as we have no idea whether it is aliased by
-some untrusted program. Because of this assumption, we also require malloc to
-promise that the non-argument registers of the register-file remain unchanged
-after the call. This allows the caller to keep private capabilities in the
-register file (strictly speaking this is not necessary as we can use our calling
-convention to protect our local state).
+Furthermore, we want to express that malloc does not depend on any other memory
+than its own internal state. This is expressed by a restriction on the malloc
+region's state interpretation function, which (as explained in
+\sectionname~\ref{subsec:worlds}) defines what memory segments it permits in a
+given world. We require that the accepted memory segments only depend on the
+presence of the malloc region itself, i.e.\ that in any world, the same memory
+segments are accepted if we remove all regions except the malloc region. This
+property should continue to hold throughout execution, so it must hold true for
+any private future region of the initial malloc region.
+
+The malloc specification then dictates what malloc should do when invoked in a
+memory with its internal state valid according to the malloc region (some future
+evolution of the initial malloc region). Malloc is invoked with a length
+argument. If this argument is invalid, that is, a negative integer or a
+capability, we simply require malloc to fail. This part of the specification
+does not actually rely on the malloc region: for simplicity, we assume malloc
+does not need its internal state to check the argument. When malloc is invoked
+with a valid length, then it should return a fresh memory segment of the correct
+length. This segment is required to come from the footprint of malloc, i.e.\ the
+memory owned by the malloc region before the call. After malloc returns, we
+require the malloc region to have evolved (according to the public future world
+relation) to a new state where the new footprint is disjoint from the allocated
+memory. This implies that future invocations of malloc can never return
+previously-allocated memory.
+
+For convenience, we also require that malloc returns the non-argument registers
+of the register-file unchanged after the call. This allows the caller to keep
+private capabilities in the register file, without having to protect them by
+storing them in a private stack frame.
 
@@
   \label{spec:malloc}
-  $c_\malloc$ satisfies the specification for malloc iff
+  $c_\malloc$ satisfies the specification for malloc iff the following conditions hold:
+  \begin{itemize}
+  \item $c_\malloc = ((\entry,\glob),\_,\_,\_)$
+  \item There exists a $\iota_{\malloc,0}$ such that
+    \begin{itemize}
+    \item $\iota_{\malloc,0}.v = \perma$
+    \item For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with $W(i)=\iota'$, we have that
+      \begin{equation*}
+        \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
+      \end{equation*}
+    \item  
+        For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$, $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
       \[
       \begin{array}{l}
-      c_\malloc = ((\entry,\glob),\_,\_,\_) \land\\
-      \exists \iota_{\malloc,0} \ldotp \\
-      \quad \iota_{\malloc,0}.v = \perma \; \land \\
-      \quad \left(
-        \begin{array}{l}
-          \forall \iota' \futurestr \iota_{\malloc,0}, W,i \ldotp\\
-          \quad W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
-        \end{array}
-        \right) \wedge \\
-      \quad \left(
-      \begin{array}{l}
-        \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
-        \quad \forall i, n, \size \in \nats \ldotp \forall
-        w_{\var{ret}} \in \Words \ldotp \\
-        \quad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
-        \quad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
-        \qquad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
-        \qquad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
-        \qquad \Rightarrow \\
-        \qquad \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-        \qquad \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
-        \qquad \qquad \quad \Phi \step[j] \Phi' \land \\
-        \qquad \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-        \qquad \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
-        \qquad \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
-        \qquad \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-        \qquad \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-        \qquad \qquad \quad \size - 1 = e'-b'
-      \end{array}
-      \right) \land\\
-      \quad \left(
-      \begin{array}{l}
-        \forall \Phi \in \ExecConfs \ldotp\\
-        \quad(\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
-        \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
-        \qquad \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed
-      \end{array}
-        \right)
+        \text{If }
+        \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
+        \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
+        \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
+        \text{Then, } \\
+        \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
+        \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+        \qquad \quad \Phi \step[j] \Phi' \land \\
+        \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+        \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
+        \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
+        \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
+        \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+        \qquad \quad \size - 1 = e'-b'
       \end{array}
       \]
+    \item
+      For all $\Phi \in \ExecConfs$,
+      \[  
+          \begin{array}{l}
+            \text{If } (\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
+            \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
+            \text{Then } \exists j \in \nats \ldotp \Phi \step[j] \failed
+          \end{array}
+      \]
+\end{itemize}
+\end{itemize}
 \end{definition}
@@
 Malloc should not just be available to trusted programs, but also to possibly
-malicious programs. The specification should therefore entail that the malloc
-capability is safe which would make it safe to let even malicious programs have
-access to it.
+malicious programs. This is safe as it follows from the specification that the malloc
+capability is always safe in a world with the malloc region:
 \begin{lemma}[Malloc is safe to pass to adversary]
@@
 The focus of this work is not to give a definitive specification and
-implementation of malloc which is why we do give a malloc implementation. We
+implementation of malloc which is why we do not give a malloc implementation. We
 are, however, confident that it is possible to make an implementation of malloc
@@
 structure malloc uses to keep track of free memory does not have to handle
-reclaimed memory. Second, the malloc specification does nor permit malloc to run
-out of memory and thus prevent allocation. This is possible on our simple
-capability machine because it has an infinite address space. The capability with
-an infinite range of authority would of course need to be part of malloc, but it
-could also double as the data structure that keeps track of free memory.
-\itoplassug{If we have references for malloc being difficult, we could add that
-  to the text.}
+reclaimed memory. Second, the malloc specification does not permit malloc to run
+out of memory and thus refuse allocation. This is possible on our simple
+capability machine because it has an infinite address space. An initial
+capability with an infinite range of authority would of course need to be part
+of malloc, but it could also double as the data structure that keeps track of
+free memory. \itoplassug{If we have references for malloc being difficult, we
+  could add that to the text.}
 %\itoplassug{Add the assumption that the malloc capability in examples

### 2018-05-06 toplas/paper.tex
Update TOPLS suggestion

--- 2018-05-04 toplas/paper.tex
@@
 \label{sec:reasoning}
-\itoplassug{
-  \begin{itemize}
-  \item A section with the lemmas we use for reasoning about programs
-    \begin{itemize}
-    \item lemma 58, scall works. Include description of how it handles the
-      common things in call and leaves the argumentation for callee code and
-      continuation.
-    \item lemma 59, malloc works. If this is included, we should probably also
-      include the malloc specification.
-    \item lemma 60, create closure works.
-    \item Include a description about how the FTLR is used to reason about
-      unknown code.
-    \item Mention that this is a step on the way towards having a program logic,
-      and then in the discussion discuss why a program logic for this language
-      is not what we need.
-    \end{itemize}
-  \end{itemize}
-}
+% \itoplassug{
+%   \begin{itemize}
+%   \item A section with the lemmas we use for reasoning about programs
+%     \begin{itemize}
+%     \item lemma 58, scall works. Include description of how it handles the
+%       common things in call and leaves the argumentation for callee code and
+%       continuation.
+%     \item lemma 59, malloc works. If this is included, we should probably also
+%       include the malloc specification.
+%     \item lemma 60, create closure works.
+%     \item Include a description about how the FTLR is used to reason about
+%       unknown code.
+%     \item Mention that this is a step on the way towards having a program logic,
+%       and then in the discussion discuss why a program logic for this language
+%       is not what we need.
+%     \end{itemize}
+%   \end{itemize}
+% }
+\itoplassug{add actual intro to this section.}
 

### 2018-05-06 toplas/paper.tex
typo
Updated informal proof of f1 lemma

@@
   calling convention, the callee can also obtain those guarantees. In other
-  words, the calling convention is modular and scales to scenario's with multiple
+  words, the calling convention is modular and scales to scenarios with multiple
   distrusting parties invoking each other.
@@
 configuration is safe (in the $\observations$ relation) for a world with a
-permanent region that requires the assertion flag to be 0. By an anti-reduction
-lemma, it suffices to show that the configuration is safe after some reduction
-steps. We then use a general lemma for reasoning about
-\texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
+permanent region that requires the assertion flag to be 0. By
+Lemma~\ref{lem:anti-red-obs}, it suffices to show that the configuration is safe
+after some reduction steps. We then use the \texttt{scall} lemma
+(Lemma~\ref{lem:scall-works}), by which it suffices to show that (1) the
 configuration that \texttt{\footnotesize{scall}} will jump to is safe and (2)
@@
 up is safe. We use the Fundamental Theorem to reason about the unknown
-adversarial code, but notice that the adversary capability is an enter
-capability, which the Fundamental Theorem says nothing about. Luckily the enter
-capability becomes $\exec$ after the jump and then the Fundamental Theorem
-applies.
+adversarial code as described in \sectionname~\ref{sec:reasoning}, but notice
+that the adversary capability is an enter capability, which the Fundamental
+Theorem says nothing about. Luckily the enter capability becomes $\exec$ after
+the jump and then the Fundamental Theorem applies.
 

### 2018-05-06 toplas/paper.tex
Changed malloc definition to enumerate, added to figure and made references from malloc description.

@@
   can safely be skipped on first read.
+  The malloc specification is presented in \figurename~\ref{fig:malloc-spec},
+  and in the following, we will connect the description of the specification by
+  referring to the item described.
 
@@
 access to malloc would have to give up this access when invoking untrusted
-code). The capability is assumed to have enter permission, so that malloc can
+code, \ref{item:malloc-spec:eg}). The capability is assumed to have enter
+permission (\ref{item:malloc-spec:eg}), so that malloc can
 protect its internal state even when the capability is shared with untrusted
@@
 To allow malloc implementations to rely on internal state and invariants for
-that state, we assume an initial region for malloc. The region is assumed to be
+that state, we assume an initial region for malloc (\ref{item:malloc-spec:reg}). The region is assumed to be
 permanent (since safety of the global malloc capability will depend on its
-presence).
+presence, \ref{item:malloc-spec:perm}).
 % DOMI: I don't understand what the following uncommented text means
@@
 property should continue to hold throughout execution, so it must hold true for
-any private future region of the initial malloc region.
+any private future region of the initial malloc region (\ref{item:malloc-spec:independent}).
 
@@
 argument. If this argument is invalid, that is, a negative integer or a
-capability, we simply require malloc to fail. This part of the specification
+capability, we simply require malloc to fail (\ref{item:malloc-spec:fail}). This part of the specification
 does not actually rely on the malloc region: for simplicity, we assume malloc
 does not need its internal state to check the argument. When malloc is invoked
-with a valid length, then it should return a fresh memory segment of the correct
+with a valid length (\ref{item:malloc-spec:correct-exec}), then it should return a fresh memory segment of the correct
 length. This segment is required to come from the footprint of malloc, i.e.\ the
@@
 
+\begin{figure}[htb]
+  \centering
   \begin{definition}[Malloc Specification]
     \label{spec:malloc}
-  $c_\malloc$ satisfies the specification for malloc iff the following conditions hold:
-  \begin{itemize}
-  \item $c_\malloc = ((\entry,\glob),\_,\_,\_)$
-  \item There exists a $\iota_{\malloc,0}$ such that
-    \begin{itemize}
-    \item $\iota_{\malloc,0}.v = \perma$
-    \item For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with $W(i)=\iota'$, we have that
+    $c_\malloc$ satisfies the specification for malloc iff the following
+    conditions hold:
+    \begin{enumerate}
+    \item \label{item:malloc-spec:eg} $c_\malloc = ((\entry,\glob),\_,\_,\_)$
+    \item \label{item:malloc-spec:reg} There exists a $\iota_{\malloc,0}$ such that
+      \begin{enumerate}
+      \item \label{item:malloc-spec:perm} $\iota_{\malloc,0}.v = \perma$
+      \item \label{item:malloc-spec:independent} For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with
+        $W(i)=\iota'$, we have that
         \begin{equation*}
@@
         \end{equation*}
-    \item  
-        For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$, $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
+      \item \label{item:malloc-spec:correct-exec} For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in
+        \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$,
+        $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
         \[
@@
         \]
-    \item
-      For all $\Phi \in \ExecConfs$,
+      \item \label{item:malloc-spec:fail} For all $\Phi \in \ExecConfs$,
         \[
@@
         \]
-\end{itemize}
-\end{itemize}
+      \end{enumerate}
+    \end{enumerate}
   \end{definition}
+
+  \caption{The specification of malloc.}
+  \label{fig:malloc-spec}
+\end{figure}
 As described previously, the specification of malloc ensures that malloc has no

### 2018-05-06 toplas/paper.tex
Added ref to malloc spec.

@@
   expresses standard expectations of how malloc should behave and this section
-  can safely be skipped on first read.
-  The malloc specification is presented in \figurename~\ref{fig:malloc-spec},
-  and in the following, we will connect the description of the specification by
+  can safely be skipped on first read. The malloc specification is presented in
+  \figurename~\ref{fig:malloc-spec} as Definition~\ref{spec:malloc}, and in
+  the following, we will connect the description of the specification by
   referring to the item described.

### 2018-05-07 toplas/paper.tex
Minor changes to (mainly) section 5 of the ToPLaS paper.

--- 2018-05-06 toplas/paper.tex
@@
   so rather than providing the implementation, we provide a specification that
-  malloc must satisfy. The specification is a bit technical, but really just
-  expresses standard expectations of how malloc should behave and this section
-  can safely be skipped on first read.
-  The malloc specification is presented in \figurename~\ref{fig:malloc-spec},
-  and in the following, we will connect the description of the specification by
-  referring to the item described.
+  malloc must satisfy. The specification expresses standard expectations of how
+  malloc should behave, but making it realistic requires using some of the
+  technical machinery from the logical relation. As such, this section is a bit
+  technical and can safely be skipped on first read. The malloc specification is
+  presented in \figurename~\ref{fig:malloc-spec}, and in the following, we will
+  connect the description of the specification by referring to the item
+  described.
 
@@
 
-In addition to these syntactic requirements, we also specify standard intuitive
-expectations for how malloc behaves. When malloc is invoked with a length
-argument, then it will return a capability for a fresh piece of memory of that
-size. It should be fresh in the sense that malloc has not already given out a
-capability for any part of that memory before and will not do so in the future.
-Finally, we also need to make sure that malloc behaves correctly when given a
-nonsensical length argument such as a negative integer or a capability. However,
+In addition to these syntactic requirements, we also specify standard
+expectations for how malloc behaves. Intuitively, we require that when malloc is
+invoked with a length argument, then it will return a capability for a fresh
+piece of memory of that size. It should be fresh in the sense that malloc has
+not already given out a capability for any part of that memory before and will
+not do so in the future. Also, when invoked with a nonsensical length argument
+such as a negative integer or a capability, malloc should simply fail. However,
 formulating these requirements is harder than one might expect. The problem is
-that we want to allow malloc implementations to rely on internal state and
-invariants on that state and it is reasonable to expect both this internal state
-and the invariants to change over time. To express that malloc should behave
-according to its specifications, but only if its internal state satisfies its
-current invariants in an executing system, we will use the semantic model
-defined in Section~\ref{sec:logical-relation} to express our specification.
+that a realistic implementation of malloc needs to rely on internal state that
+changes after every invocation, and relies on invariants on that state. We can
+only expect that malloc behaves according to its specification, if its internal
+state satisfies the current state of the invariants in an executing system. To
+express this, we will use the semantic model defined in
+Section~\ref{sec:logical-relation}.
 
@@
 memory with its internal state valid according to the malloc region (some future
-evolution of the initial malloc region). Malloc is invoked with a length
-argument. If this argument is invalid, that is, a negative integer or a
-capability, we simply require malloc to fail (\ref{item:malloc-spec:fail}). This part of the specification
-does not actually rely on the malloc region: for simplicity, we assume malloc
-does not need its internal state to check the argument. When malloc is invoked
-with a valid length (\ref{item:malloc-spec:correct-exec}), then it should return a fresh memory segment of the correct
-length. This segment is required to come from the footprint of malloc, i.e.\ the
-memory owned by the malloc region before the call. After malloc returns, we
-require the malloc region to have evolved (according to the public future world
-relation) to a new state where the new footprint is disjoint from the allocated
-memory. This implies that future invocations of malloc can never return
-previously-allocated memory.
+evolution of the initial malloc region). If malloc is invoked with an invalid
+length argument (that is, a negative integer or a capability), then we simply
+require malloc to fail (\ref{item:malloc-spec:fail}). This part of the
+specification does not actually rely on the malloc region: for simplicity, we
+assume malloc does not need its internal state to check the argument. When
+malloc is invoked with a valid length (\ref{item:malloc-spec:correct-exec}),
+then it should return a fresh memory segment of the correct length. This segment
+is required to come from the footprint of malloc, i.e.\ the memory owned by the
+malloc region before the call. After malloc returns, we require the malloc
+region to have evolved (according to the public future world relation) to a new
+state where the new footprint is disjoint from the allocated memory. This
+implies that future invocations of malloc can never return previously-allocated
+memory.
 
@@
 capabilities pointing \emph{out} of malloc. It does, however, not say anything about
-capabilities that point \emph{in} to malloc. We obviously can not
+capabilities that point \emph{in} to malloc. We obviously cannot
 allow this if we want to trust malloc. We have chosen to keep the malloc
@@
 
-The focus of this work is not to give a definitive specification and
-implementation of malloc which is why we do not give a malloc implementation. We
-are, however, confident that it is possible to make an implementation of malloc
-that satisfies the malloc specification in Definition~\ref{spec:malloc}. There
-are in particular two simplifications in our system that makes things easy:
-First, we do not consider deallocation of memory which means that the data
-structure malloc uses to keep track of free memory does not have to handle
-reclaimed memory. Second, the malloc specification does not permit malloc to run
-out of memory and thus refuse allocation. This is possible on our simple
-capability machine because it has an infinite address space. An initial
-capability with an infinite range of authority would of course need to be part
-of malloc, but it could also double as the data structure that keeps track of
-free memory. \itoplassug{If we have references for malloc being difficult, we
-  could add that to the text.}
+The reason we allow trusted code and adversary to invoke malloc is just to make
+our work more realistic, but we are otherwise not interested in its details. As
+such, we do not give a malloc implementation. We are, however, confident that it
+is possible to make an implementation of malloc that satisfies the malloc
+specification in Definition~\ref{spec:malloc}. There are in fact two
+simplifications in our system that makes things easier: First, we do not consider
+deallocation of memory which means that the data structure malloc uses to keep
+track of free memory does not have to handle reclaimed memory. Second, the
+malloc specification does not permit malloc to run out of memory and thus refuse
+allocation. This is possible on our simple capability machine because it has an
+infinite address space. An initial capability with an infinite range of
+authority would of course need to be part of malloc, but it could also double as
+the data structure that keeps track of free memory. % \itoplassug{If we have
+  % references for malloc being difficult, we could add that to the text.}
 %\itoplassug{Add the assumption that the malloc capability in examples
@@
 I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
-With the semantic model defined, we would like to show its usefulness by proving
+With the calling convention and logical relation defined, we would like to show its usefulness by proving
 the correctness of a series of examples that rely
@@
 convention to enforce these properties. However, the programs that
-run on our capability are assembly programs and even program examples that would
+run on our capability machine are assembly programs and even program examples that would
 be small in a high-level language become big and unintelligible at this
@@
 first thing upon return, and restores the stack pointer and the old program
-pointer. After the restoration code is pushed to the stack, a copy of the pc is
+counter. After the restoration code is pushed to the stack, a copy of the pc is
 pushed to the stack after it has been adjusted to point to the first instruction
@@
 adjusting it to point to the return code and encapsulating it by restricting its
-permission to Enter. Next, the stack pointer is restricted to the unused part,
-and the unused part of the stack is cleared, so it does not govern any part of
-the private stack frame which means that it can safely be given away. Finally,
-the non-argument registers are cleared and we jump to $r$. Upon return after the
-restoration code has been executed, the restoration code is popped from the
-stack and the private words we pushed to the stack before the call are popped to
-the private registers. The implementation of \texttt{scall} is presented in
+permission to $\entry$. Next, the stack pointer is restricted to the unused
+part, and the unused part of the stack is cleared (as discussed in
+Section~\ref{sec:stack-and-return-pointer}). Finally, the non-argument registers
+are cleared and we jump to $r$. Upon return after the restoration code has been
+executed, the restoration code is popped from the stack and the private words we
+pushed to the stack before the call are popped to the private registers. The
+implementation of \texttt{scall} is presented in
 \figurename~\ref{fig:scall-impl}, and the restoration code is presented in
 \figurename~\ref{fig:scall-rest-code}. The implementation of \texttt{scall} uses
-some of the macros we now present.
+some of the macros we present next.
 
@@
 follows the convention for the stack\footnote{The stack capability should always
-point to the top most word on the stack. A stack received from an unknown source
+point to the topmost word on the stack. A stack received from an unknown source
 can be treated as empty, so the stack capability should point just outside its
@@
 contents of $\overline{r_i}$ is stored in the environment. In the code referred
-to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
+to by $r$, an implicit load from the environment happens when an instruction
+refers to $x_i$.
 
@@
 % Anti-reduction lemma
-Out capability machine only allows the final memory of an execution to be
+Our capability machine only allows the final memory of an execution to be
 observed, so naturally the correctness lemmas we prove in
 \sectionname~\ref{sec:examples} are statements about the memory in the halted
-configuration. In order to prove something about the final memory, we encode it
-as a region in a world and show that the initial configuration is in the
-$\observations$-relation w.r.t.\ that world. The $\observations$-relation says
-that if the initial configuration halts successfully, then the memory in the
-halting state must still respect a private future world of the initial world. It
-is, however, bothersome and error prone to try to argue about the entire
-execution in one go. Instead we want to reason modularly, in the sense that we
-only want to reason about parts of the execution at a time. To this end, we
-prove an anti-reduction lemma that essentially says that if we can show that
-an initial configuration, $\Phi$, steps to a configuration $\Phi'$ and
-that $\Phi'$ is in the $\observations$-relation, then $\Phi$ is also in the
-observation relation.
+configuration. In order to prove something about some part of the final memory,
+we establish an invariant for it as a region in a world and show that the
+initial configuration is in the $\observations$-relation w.r.t.\ that world. The
+$\observations$-relation says that if the initial configuration halts
+successfully, then the memory in the halting state must still respect a private
+future world of the initial world. It is, however, bothersome and error prone to
+try to argue about the entire execution in one go. Instead we want to reason
+modularly, in the sense that we only want to reason about parts of the execution
+at a time. To this end, we prove an anti-reduction lemma that essentially says
+that if we can show that an initial configuration, $\Phi$, steps to a
+configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then
+$\Phi$ is also in the observation relation.
 \begin{lemma}[Anti-reduction for $\observations$]
@@
 with a malloc capability and otherwise consists of instructions. This allows us
-to use the FTLR seeing as Malloc is a safe value cf.
-Lemma~\ref{lem:malloc-in-vr} and instructions are really just integers and they
+to use the FTLR seeing as malloc is a safe value (cf.
+Lemma~\ref{lem:malloc-in-vr}) and instructions are really just integers and they
 are always safe values. Here, it is important to remember that the FTLR gives us
@@
 \end{lemma}
-The essence of the \texttt{scall} lemma is to show \texttt{scall} is executed in
-a reasonable state (\ref{item:scall:memsat}-\ref{item:scall:stack}), it is safe
-to actually transfer control to the callee (\ref{item:scall:hyp-callee}), and that
-the return capability given to the callee is safe (\ref{item:scall:hyp-cont}),
-which corresponds to showing that the remainder of the program is safe to execute.
+Roughly, the \texttt{scall} lemma states that an invocation of scall is safe if
+\texttt{scall} is executed in a reasonable state
+(\ref{item:scall:memsat}-\ref{item:scall:stack}), the callee is safe to execute
+(\ref{item:scall:hyp-callee}), and returning to the code after the scall is safe
+(\ref{item:scall:hyp-cont}).
 

### 2018-05-07 toplas/paper.tex
Added some more text about the future region relations and in particular masks

--- 2018-05-06 toplas/paper.tex
@@
 to transition according to the private part of the transition system.
+
+\begin{toplas}
+  In the future world relations, we must remember all region names in order to
+  keep them extensional which is why we need to use masks. It does, however, not
+  matter what kind of region is used to mask as it is the region name we must
+  retain. This is why we allow $\revoked$ regions to be replaced by other
+  regions in the two future world relations. This may seem unnecessary as the
+  future world relation allows new regions to be added, but in private future
+  worlds, for technical reasons, we need to be able to reinstate old regions
+  with the same region names as they had in a past world. Further, we do not
+  want public future worlds to be able to prevent reinstation in private future
+  worlds. Private future worlds cannot revoke $\perma$ regions, so we cannot
+  allow public future worlds to us $\perma$ regions as a mask for a revoked
+  region.
+\end{toplas}
 

### 2018-05-07 toplas/paper.tex
Minor changes to section 7.

--- 2018-05-07 toplas/paper.tex
@@
 In the common case \texttt{scall} is used to invoke untrusted code, and we will
-only have very general assumptions about that code such as it has contains a
-capability for a linking table with malloc in it and all other addresses are
-instructions. These values are all safe which allows us to conclude that the
-read condition holds for the untrusted code which allows us to use the
-Theorem~\ref{thm:ftlr} to conclude that it can be safely executed if the memory
-and the register-file are safe. By assumption in Hyp-Callee, the memory is safe,
-so it suffices to show that the register-file contains safe values, which
-amounts to showing that the arguments in the call are safe.
+only have very general assumptions about that code, in which case we will use
+the FTLR (Theorem~\ref{thm:ftlr}) to argue
+Assumption~\ref{item:scall:hyp-callee}, using what we know about the values it
+has access to (e.g.\ linking table, malloc capability etc.). By assumption in
+Hyp-Callee, the memory is safe, so it suffices to show that the register-file
+contains safe values, which amounts to showing that the arguments in the call
+are safe.
 
@@
 (\ref{item:malloc:looking-at}-\ref{item:malloc:malloc-sat}) as well as
-requirements on the execution afterwards (\ref{item:malloc:hyp-cont}). It does,
-however, not have any requirements on the callee as we can rely on the
-\texttt{malloc} specification to reason about this part once and for all.
+requirements on the execution afterwards (\ref{item:malloc:hyp-cont}). It does
+not have any requirements on the callee as the \texttt{malloc} specification
+defines its behavior.
 \begin{lemma}[$\mathtt{malloc}$ works]
@@
 In the technical appendix~\citep{technical_appendix}, we also define a lemma for
-the macro used to create closures, \texttt{crtcls}. We stress that these lemmas
-are conveniences and are not strictly necessary to reason about programs for the
-capability machine. In principle, one could reason solely in the semantic model
-without the use of these lemmas, but that would lead to a lot of repetition.
+the macro used to create closures, \texttt{crtcls}.
+% The below remark seems unnecessary: isn't this true about any lemma in any proof?
+% We stress that these lemmas
+% are conveniences and are not strictly necessary to reason about programs for the
+% capability machine. In principle, one could reason solely in the semantic model
+% without the use of these lemmas, but that would lead to a lot of repetition.
 \end{toplas}

### 2018-05-07 toplas/paper.tex
Add an intro to section 7 + some other minor changes.

--- 2018-05-07 toplas/paper.tex
@@
 % }
-\itoplassug{add actual intro to this section.}
+There are many details to get right when programming in assembly. These details
+carry over to proofs about assembly programs, so many of the proofs in the next
+section about example programs are a bit cumbersome. It is especially annoying,
+when the same line of reasoning is applied in multiple places. To mitigate this,
+we have defined a number of lemmas that capture common reasoning patterns in
+these proofs. We present the most important such lemmas in this section, to give
+the reader an idea of how they work.
 
@@
 % Many details to get right when reasoning about low-level programs.
-There are many details to get right when programming in assembly. These details
-carry over to proofs about assembly programs, so the proofs become cumbersome.
-It is especially annoying, when the same line of reasoning is applied in
-multiple places. For instance, in $\texttt{scall}$ the setup before the jump to
-the call is always the same and the recovery upon return is also always the
-same, and in both cases the reasoning will be the same every time.
-We deal with this part of the reasoning in the following lemma.
+Another common pattern in proofs about programs on our capability machine
+concerns the use of scall. The following lemma captures the commonalities of
+reasoning about programs using scall. From another point of view, it can be seen
+as a specification for scall.
 \begin{lemma}[$\mathtt{scall}$ works]
@@
 
-The \texttt{malloc} instruction is also used often, so we also prove
+Also often used is the \texttt{malloc} instruction, so we also prove
 Lemma~\ref{lem:malloc-works} to help reason about it. The structure of the

### 2018-05-07 toplas/paper.tex
Make appendix lemmas refer to the originals in the paper.

@@
  \subsection{Example correctness lemmas}
-\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
+\begin{lemma}[Correctness lemma for \texttt{f1}, copy of Lemma~\ref{lem:correctness-f1}] \forcenewline
   \label{lem:correctness-f1-app}
@@
 
-\begin{lemma}[Correctness lemma for \texttt{f2}]
+\begin{lemma}[Correctness lemma for \texttt{f2}, detailed version of Lemma~\ref{lem:correctness-f2}]
   \label{lem:correctness-f2-app}
@@
 \end{lemma}
-\begin{lemma}[Correctness lemma for \texttt{f3}]
-  \label{lem:correctness-f3}
+\begin{lemma}[Correctness lemma for \texttt{f3}, detailed version of Lemma~\ref{lem:correctness-f3}]
+  \label{lem:correctness-f3-detailed}
   For all $n \in \nats$
@@
 
- \begin{lemma}[Correctness of $g1$]
-  \label{lem:correctness-g1}
+ \begin{lemma}[Correctness of $g1$, detailed version of Lemma~\ref{lem:correctness-g1}]
+  \label{lem:correctness-g1-detailed}
   For all $n \in \nats$

### 2018-05-07 toplas/paper.tex
Additional note in related work section.
Small changes to different parts of the paper.
reformulations
added two definitions to the appendix

--- 2018-05-07 toplas/paper.tex
@@
 \begin{toplas}
-\todo[inline]{Look this introduction and the following section through to check for overlap.
+%\todo[inline]{Look this introduction and the following section through to check for overlap.
 Also consider making backwards references to this introduction from the places where we use advanced techniques, so it becomes clear that they are necessary.}
@@
 reasons, we omit some definitions and explain them only verbally, but
-precise definitions can be found in the technical
-appendix~\citep{technical_appendix}.
+precise definitions can be found in the appendix.
 
@@
 program counter (in the $\stder$-relation) no matter the current address.
-\itoplassug{The range of an executable capability can also be updated, so they must also be safe as program counter no matter what their range of authority is reduced to.}
+\itoplas{The range of an executable capability can also be reduced, so they must also be safe as program counter no matter what their range of authority is reduced to.}
 In contrast, enter capabilities are opaque and can only be used to jump to the
@@
 % }
-There are many details to get right when programming in assembly. These details
-carry over to proofs about assembly programs, so many of the proofs in the next
-section about example programs are a bit cumbersome. It is especially annoying,
-when the same line of reasoning is applied in multiple places. To mitigate this,
-we have defined a number of lemmas that capture common reasoning patterns in
-these proofs. We present the most important such lemmas in this section, to give
-the reader an idea of how they work.
+\itoplassug{add actual intro to this section.}
 
@@
 % Many details to get right when reasoning about low-level programs.
-Another common pattern in proofs about programs on our capability machine
-concerns the use of scall. The following lemma captures the commonalities of
-reasoning about programs using scall. From another point of view, it can be seen
-as a specification for scall.
+There are many details to get right when programming in assembly. These details
+carry over to proofs about assembly programs, so the proofs become cumbersome.
+It is especially annoying, when the same line of reasoning is applied in
+multiple places. For instance, in $\texttt{scall}$ the setup before the jump to
+the call is always the same and the recovery upon return is also always the
+same, and in both cases the reasoning will be the same every time.
+We deal with this part of the reasoning in the following lemma.
 \begin{lemma}[$\mathtt{scall}$ works]
@@
 
-Also often used is the \texttt{malloc} instruction, so we also prove
+The \texttt{malloc} instruction is also used often, so we also prove
 Lemma~\ref{lem:malloc-works} to help reason about it. The structure of the
@@
   \activeReg{W} & \defeq \dom(\erase{W}{\perma,\temp})
+\end{align*}
+\textbf{Projection of regions based on locality}
+\[
+  \localityReg(\gl,W) \defeq 
+  \begin{cases}
+    \dom(\erase{W}{\perm,\temp}) & \text{if } \gl = \local \\
+    \dom(\erase{W}{\perm}) & \text{if } \gl = \glob
+  \end{cases}
+\]
+\textbf{Address stratification}
+\begin{align*}
+&\iota = (v,s,\phi_\pub,\phi,H) \text{ is address-stratified iff }\\
+&\qquad\begin{multlined}
+  \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
+  \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
+  \dom(\ms) = \dom(\ms') \wedge \\
+  \forall \addr \in
+  \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
+\end{multlined}
 \end{align*}

### 2018-05-07 toplas/paper.tex

--- 2018-05-07 toplas/paper.tex
@@
 \begin{toplas}
-In the future we would also like to
-investigate a new program logic for reasoning about capability safety for our
-capability machine model. We think Iris would also be a natural starting point
-for such an endeavour, since Iris is really a framework, which can be
-instantiated to different programming languages. OCPL was able to leverage
-existing Iris specifications for the high-level language; for our capability
-machine model, however, it would be necessary to devise new kinds of
-specifications for our low-level programs with unstructured control-flow. It is
-likely that we could get inspiration from earlier work on logics for assembly
-programming languages, such as XCAP~\citep{xcap}.
+In the future we would also like to investigate a new program logic for
+reasoning about capability safety for our capability machine model. In fact,
+we think the lemmas in Section~\ref{sec:reasoning} are suggestive of the style
+of results that could be written in such a logic. We think Iris would also be
+a natural starting point for such an endeavour, since Iris is really a
+framework, which can be instantiated to different programming languages. OCPL
+was able to leverage existing Iris specifications for the high-level language;
+for our capability machine model, however, it would be necessary to devise new
+kinds of specifications for our low-level programs with unstructured
+control-flow. It is likely that we could get inspiration from earlier work on
+logics for assembly programming languages, such as XCAP~\citep{xcap}.
 \end{toplas}

### 2018-05-07 toplas/paper.tex
Add short introductory paragraph to appendix.

--- 2018-05-07 toplas/paper.tex
@@
 \section{Appendix}
+In this appendix, we give more precise formulations of lemmas that were mentioned in the paper, and the most important supporting definitions and lemmas.
+The goal is to provide details that can help to understand in more detail what we discuss in the paper. 
+Full details and proofs are not given here, but for those we refer to the technical appendix~\citep{technical_appendix}.
+
 \subsection{Logical relation}

### 2018-05-08 toplas/paper.tex
appendix, cofes

--- 2018-05-07 toplas/paper.tex
@@
 % Uncomment me to disable comments
-%\renewcommand\lau[1]{}
-%\renewcommand\dominique[1]{}
-%\renewcommand\lars[1]{}
+\renewcommand\lau[1]{}
+\renewcommand\dominique[1]{}
+\renewcommand\lars[1]{}
 
@@
 % Conv defs
-\newcommand{\linksto}[4]{\ensuremath{#1} \text{ links } \ensuremath{#2} \text{ as } \ensuremath{#3} \text{ to } \ensuremath{#4}}
 \newcommand{\lookingat}[3]{\ensuremath{#1} \text{ is looking at } \ensuremath{#2} \text{ followed by } \ensuremath{#3}}
@@
 
-\dominique{Follow order of figure in discussion?}
 The $\load{}{}$ instruction loads the contents of the requested memory location
@@
 in software, precisely to allow experimenting with alternative models like ours.
-\dominique{Update to CHERI's new ccall story.}
 
@@
 further in the next section.
-\itoplas{The specification of malloc is presented in \sectionname~\ref{sec:malloc} as it uses the semantic model we build in Section~\ref{sec:logical-relation}.
+\itoplas{The specification of malloc is presented in \sectionname~\ref{sec:malloc-and-macros} as it uses the semantic model we build in Section~\ref{sec:logical-relation}.
 For full details on the linking table, we refer to the technical
@@
   linking model)\\}
-\itoplassug{Suggestion: We also have the flag table used
+\itoplassug{Suggestion: We also have the flag table ussed
   for assertions, maybe we should add a description of it here.\\}
@@
   the callee could be well-behaved and wish to ensure well-bracketed calls and
-  local-state encapsulation as well. The calling convention puts no restriction
-  on the callee that the caller itself does not follow, so by following the
-  calling convention, the callee can also obtain those guarantees. In other
-  words, the calling convention is modular and scales to scenarios with multiple
-  distrusting parties invoking each other.
+  local-state encapsulation. The calling convention puts no restriction on the
+  callee that the caller itself does not follow, so by following the calling
+  convention, the callee would also get well-bracketedness and local-state
+  encapsulation. Generally, the calling convention ensures the two properties
+  for anyone who employs it, so it scales to the setting where multiple
+  distrusting parties need to invoke each other and each of them can protect
+  themself by using the calling convention.
 \end{toplas}
@@
 to transition according to the private part of the transition system.
-
-\begin{toplas}
-  In the future world relations, we must remember all region names in order to
-  keep them extensional which is why we need to use masks. It does, however, not
-  matter what kind of region is used to mask as it is the region name we must
-  retain. This is why we allow $\revoked$ regions to be replaced by other
-  regions in the two future world relations. This may seem unnecessary as the
-  future world relation allows new regions to be added, but in private future
-  worlds, for technical reasons, we need to be able to reinstate old regions
-  with the same region names as they had in a past world. Further, we do not
-  want public future worlds to be able to prevent reinstation in private future
-  worlds. Private future worlds cannot revoke $\perma$ regions, so we cannot
-  allow public future worlds to us $\perma$ regions as a mask for a revoked
-  region.
-\end{toplas}
 
@@
 In contrast, enter capabilities are opaque and can only be used to jump to the
-address they point to.%
-\begin{toplas}%
-  This is why the $\entryCond{}$ depends on the current address of the capability, unlike for other types of capabilities.
+address they point to.
+\begin{toplas}
+  As enter capabilities are the only capabilities that have to use the address they point to, it is natural that the $\entryCond{}$ is also the only permission-based condition that depends on the current address of the capability.
 \end{toplas}
@@
 \begin{toplas}
-  \begin{proof}[Proof sketch]
+  \begin{proof}
     Induction over $n$. By definition of $\stder(W)$, show
     $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
-    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
+    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg}$, and
     $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
-    $\mem'$, and $i \leq n'$ be given and for
-    $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms \uplus
-    \ms_f)$ assume $\Phi \step[i]
+    $\mem'$, and $i \leq n'$ be given and assume $\Phi \step[i]
     (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
-    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halted
+    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halting
     configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
-    \Phi' \step[i-1] (\halted,\mem')$.
-
-    The rest of the proof considers the different ways the step $\Phi \step
-    \Phi'$ could have occurred, depending on the instruction being executed. For
-    each of these cases, we argue (1) that $\Phi'$ is consistent with the world
-    in the sense that the register-file and memory still respect the world and
-    (2) that the rest of the execution respects the world. Depending on where
-    the $\pcreg$ in $\Phi'$ comes from, the second result is proven in one of
-    two ways. If the step to $\Phi'$ was a jump, then the new $\pcreg$ is one of
-    the safe values in $\Phi$'s registers and the value relation can be used to
-    argue that those can be safely jumped to. On the other hand, if the $\pcreg$
-    was just incremented, then we can apply the induction hypothesis.
-
-    In order to argue (1), we consider what configurations the operational
-    semantics allows us to get to from the initial state. If we consider the
-    memory in $\Phi'$, then it either (a) remains unchanged or (b) one address
-    has been updated and in that case, the register-file contains an appropriate
-    capability for writing. The latter occurs when the instruction being
-    executed is a $\texttt{store}$, and otherwise the former. In case (a), we
-    can conclude that the memory still respects the world just by downwards
-    closure of memory satisfaction. In case (b), we use an auxiliary lemma that
-    uses the safety of the write capability used by the $\texttt{store}$
-    instruction to show that the updated memory satisfies the world. To show
-    that the updated register-file is safe, we consider the changes made to it
-    by all instructions in separate lemmas and show that they all preserve
-    safety of the register file.
+    \var{conf}' \step[i-1] (\halted,\mem')$.
+
+    The rest of the proof roughly consists of two parts (1) argue that
+    $\var{conf}'$ is consistent with the world in the sense that the
+    register-file and memory still respect the world. (2) Argue that the rest of
+    the execution respects the world. The second case, relies on where the
+    $\pcreg$ in $\var{conf}'$ comes from. If the step to $\var{conf}'$ was a
+    jump, then use the fact that the register-file contains safe values in the
+    initial configuration. On the other hand, if the $\pcreg$ was just
+    incremented, then we need to apply the induction hypothesis which is
+    possible as all of the permission based conditions are downwards closed.
+    For instance in the case where the $\pcreg$ is read-execute, i.e.\ $\perm =
+    \exec$, we also assume $\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)$
+    which entails $\npair[n'-1]{(\start,\addrend)} \in \readCond{}(\gl)(W)$ and
+    allows us to use the induction hypothesis.
+    In order to argue (1), we just need to consider the configurations the
+    operational semantics allows us to get to from the initial state. If we
+    consider the memory in $\var{conf}'$, then it either (a) remains unchanged
+    or (b) one address has been updated and in that case, the register-file contains an 
+    appropriate capability for writing. The latter case, (b) corresponds to the first
+    step executing a $\texttt{store}$ instruction, and the former (a) corresponds to
+    any other instruction.
+    In case (a), we can conclude that the memory still respects the world just
+    by downwards closure of memory satisfaction. In the other case, (b), we use
+    an auxiliary lemma called \emph{Conditions for \texttt{store} are
+      sufficient} (see the appendix, Lemma~\ref{lem:conds-store-stuff})
+    that states that given all the assumptions we have done so far, the updated
+    memory still satisfies the world.
+    All the possible changes to the register-file correspond to the effect of
+    some instruction, so for the non-trivial instructions we use a sufficiency
+    lemma.
 
@@
 \begin{toplas}
-  \section{Malloc}
-  \label{sec:malloc}
-  In the examples we present in \sectionname~\ref{sec:examples}, we will assume
-  the existence of a trusted malloc routine so that both the trusted code and
-  the adversary can allocate new memory. Malloc is considered part of the
-  trusted computing base, as mentioned in
-  \sectionname~\ref{sec:capab-mach-with}. This is unavoidable: if we cannot
-  trust malloc, then we cannot use the memory it allocates as we have no idea
-  whether it is aliased by some untrusted program.
-
-  Our semantic model is not specific to a particular implementation of malloc,
-  so rather than providing the implementation, we provide a specification that
-  malloc must satisfy. The specification expresses standard expectations of how
-  malloc should behave, but making it realistic requires using some of the
-  technical machinery from the logical relation. As such, this section is a bit
-  technical and can safely be skipped on first read. The malloc specification is
-  presented in \figurename~\ref{fig:malloc-spec}, and in the following, we will
-  connect the description of the specification by referring to the item
-  described.
-
-We require a global capability for
-invoking malloc (because if the capability were local, then a program with
-access to malloc would have to give up this access when invoking untrusted
-code, \ref{item:malloc-spec:eg}). The capability is assumed to have enter
-permission (\ref{item:malloc-spec:eg}), so that malloc can
-protect its internal state even when the capability is shared with untrusted
-code.
-
-In addition to these syntactic requirements, we also specify standard
-expectations for how malloc behaves. Intuitively, we require that when malloc is
-invoked with a length argument, then it will return a capability for a fresh
-piece of memory of that size. It should be fresh in the sense that malloc has
-not already given out a capability for any part of that memory before and will
-not do so in the future. Also, when invoked with a nonsensical length argument
-such as a negative integer or a capability, malloc should simply fail. However,
-formulating these requirements is harder than one might expect. The problem is
-that a realistic implementation of malloc needs to rely on internal state that
-changes after every invocation, and relies on invariants on that state. We can
-only expect that malloc behaves according to its specification, if its internal
-state satisfies the current state of the invariants in an executing system. To
-express this, we will use the semantic model defined in
-Section~\ref{sec:logical-relation}.
-
-To allow malloc implementations to rely on internal state and invariants for
-that state, we assume an initial region for malloc (\ref{item:malloc-spec:reg}). The region is assumed to be
-permanent (since safety of the global malloc capability will depend on its
-presence, \ref{item:malloc-spec:perm}).
-% DOMI: I don't understand what the following uncommented text means
-% This also means that the malloc specification should guarantee that
-% use of malloc is safe at any point of execution no matter who has invoked it. To
-% this end, the requirement on the initial malloc region should also hold true in
-% any private future world.
+  \section{Malloc and macros}
+  \label{sec:malloc-and-macros}
+With the semantic model defined, we would like to show its usefulness by proving
+the correctness of a series of examples that rely
+on well-bracketedness and local-state encapsulation and use the calling
+convention to enforce these properties. However, the programs that
+run on our capability are assembly programs and even program examples that would
+be small in a high-level language, they become big and unintelligible at this
+low level. In order to make our program examples intelligible, we introduce a
+series of low-level abstractions in the form of macros.
+
+To make the program examples realistic, they have access to a trusted malloc, so
+they can allocate new memory. Our results are not specific to
+any one implementation of malloc, so rather than providing the implementation,
+we provide a specification that malloc must satisfy.
+
+\subsection{Malloc specification}
+  \itoplassug{Suggestion: add the malloc specification}
+Before we give the malloc specification, we provide some intuition about its requirements.
+
+Malloc is assumed to be part of the trusted computing base, as mentioned in
+\sectionname~\ref{sec:capab-mach-with}. A program with access to malloc should
+be able to retain this access even if it invokes code it does not trust. In
+order to achieve this, a capability for malloc must be global. It is also
+important that malloc remains trustworthy to all entities on the machine
+throughout execution. That is, malloc should encapsulate its local state and
+not depend on any external code. To make sure that malloc cannot be tampered
+with directly through the malloc capability, we require it to have enter
+permission.
+
+The requirements so far are syntactic and prevent direct tampering in the sense
+that the capability intended for invoking malloc only can be used for
+invocation. This does, however, not prevent the implementation of malloc from
+colluding with other code. Specifically, malloc could depend on the state of a
+piece of memory not just under the control of malloc. We also need to make sure
+that malloc actually behave as intended. That is when malloc is invoked with a
+length argument, then it actually returns a capability for a fresh piece of
+memory of that size. It should be fresh in the sense that malloc hasn't given
+out a capability for any part of that memory before and won't do it in the
+future. Finally, we also need to make sure that malloc behaves correctly when
+given a nonsensical length argument such as a negative integer or a capability.
+These requirements relate to the behaviour of malloc, so we cannot hope to
+express them syntactically. Instead we will use the semantic model defined in
+Section~\ref{sec:logical-relation} to express this part of the specification.
+
+In order to express the semantic requirements on malloc, the specification will
+require the existence of some initial region for malloc. Malloc does not
+disappear from the system, so the malloc region must be permanent. This also
+means that the malloc specification should guarantee that use of malloc is safe
+at any point of execution no matter who has invoked it. To this end, the
+requirement on the initial malloc region should also hold true in any private
+future world.
 % priv future means that malloc should not depend on local capabilities?
-Furthermore, we want to express that malloc does not depend on any other memory
-than its own internal state. This is expressed by a restriction on the malloc
-region's state interpretation function, which (as explained in
-\sectionname~\ref{subsec:worlds}) defines what memory segments it permits in a
-given world. We require that the accepted memory segments only depend on the
-presence of the malloc region itself, i.e.\ that in any world, the same memory
-segments are accepted if we remove all regions except the malloc region. This
-property should continue to hold throughout execution, so it must hold true for
-any private future region of the initial malloc region (\ref{item:malloc-spec:independent}).
-
-The malloc specification then dictates what malloc should do when invoked in a
-memory with its internal state valid according to the malloc region (some future
-evolution of the initial malloc region). If malloc is invoked with an invalid
-length argument (that is, a negative integer or a capability), then we simply
-require malloc to fail (\ref{item:malloc-spec:fail}). This part of the
-specification does not actually rely on the malloc region: for simplicity, we
-assume malloc does not need its internal state to check the argument. When
-malloc is invoked with a valid length (\ref{item:malloc-spec:correct-exec}),
-then it should return a fresh memory segment of the correct length. This segment
-is required to come from the footprint of malloc, i.e.\ the memory owned by the
-malloc region before the call. After malloc returns, we require the malloc
-region to have evolved (according to the public future world relation) to a new
-state where the new footprint is disjoint from the allocated memory. This
-implies that future invocations of malloc can never return previously-allocated
-memory.
-
-For convenience, we also require that malloc returns the non-argument registers
-of the register-file unchanged after the call. This allows the caller to keep
-private capabilities in the register file, without having to protect them by
-storing them in a private stack frame.
-
-\begin{figure}[htb]
-  \centering
+
+The implementation of malloc should not depend on any memory outside the
+implementation of malloc. This means that malloc should not have any
+capabilities in its internals that govern other programs. As explained in
+\sectionname~\ref{subsec:worlds}, a region's state interpretation function
+determines what memory segments it permits with respect to some world. The state
+interpretation function of malloc should permit the same regions with respect to
+a world with the malloc region as it would with respect to a world that only
+contains the malloc region. In other words, the remaining regions in the world
+do not influence which regions malloc accepts. Malloc should retain this
+property throughout execution, so it must hold true for any private future
+region of the initial malloc region.
+
+The malloc specification will also dictate what malloc should do when invoked.
+Malloc can be invoked with either a valid or an invalid length argument. In the
+case where malloc is invoked with an invalid argument, that is a negative
+integer or a capability, we require malloc to simply fail. This part of the
+specification does not rely on the malloc region as failing is considered
+safe behaviour. When malloc is invoked with a valid length, then malloc should
+return a fresh memory segment of the correct length.
+\todo[inline]{Everything below this is work in progress..}
+Malloc may have been invoked a number of times before any given invocation, so
+the internal data structures of malloc may be in a different state than it
+started in.
+We know that the memory of malloc is in a state consistent with the
+protocol dictated by the initial malloc region in the sense that we can assume
+that malloc is invoked in a configuration where part of it satisfies a private
+future region of the initial malloc region. Knowing the state of the malloc
+region allows us to express what we mean by a \emph{fresh} region. This means
+that malloc should not keep any capabilities for the freshly allocated region.
+To ensure this, we require the footprint of malloc after the call to be disjoint
+from the freshly allocated memory. Further, the footprint should satisfy a
+public future region of the malloc region the invocation happened in. This means
+that it cannot depend on the freshly allocated region. It further allows
+subsequent calls to malloc as it provides one of the requirements in the malloc
+specification.
+
+Malloc is part of the trusted computing base. If we cannot trust malloc, then we
+cannot use the memory it allocates as we have no idea whether it is aliased by
+some untrusted program. Because of this assumption, we also require malloc to
+promise that the non-argument registers of the register-file remain unchanged
+after the call. This allows the caller to keep private capabilities in the
+register file. This is strictly speaking not necessary as we can use our calling
+convention to protect our local state.
+
 \begin{definition}[Malloc Specification]
   \label{spec:malloc}
-    $c_\malloc$ satisfies the specification for malloc iff the following
-    conditions hold:
-    \begin{enumerate}
-    \item \label{item:malloc-spec:eg} $c_\malloc = ((\entry,\glob),\_,\_,\_)$
-    \item \label{item:malloc-spec:reg} There exists a $\iota_{\malloc,0}$ such that
-      \begin{enumerate}
-      \item \label{item:malloc-spec:perm} $\iota_{\malloc,0}.v = \perma$
-      \item \label{item:malloc-spec:independent} For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with
-        $W(i)=\iota'$, we have that
-        \begin{equation*}
-          \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
-        \end{equation*}
-      \item \label{item:malloc-spec:correct-exec} For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in
-        \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$,
-        $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
+  $c_\malloc$ satisfies the specification for malloc iff
   \[  
     \begin{array}{l}
-            \text{If }
-            \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
-            \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
-            \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
-            \text{Then, } \\
-            \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-            \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
-            \qquad \quad \Phi \step[j] \Phi' \land \\
-            \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-            \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
-            \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
-            \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-            \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
-            \qquad \quad \size - 1 = e'-b'
+      c_\malloc = ((\entry,\glob),\_,\_,\_) \land\\
+      \exists \iota_{\malloc,0} \ldotp \\
+      \quad \iota_{\malloc,0}.v = \perma \; \land \\
+      \quad \left(
+        \begin{array}{l}
+          \forall \iota' \futurestr \iota_{\malloc,0}, W,i \ldotp\\
+          \quad W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
+        \end{array}
+        \right) \wedge \\
+      \quad \left(
+      \begin{array}{l}
+        \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
+        \quad \forall i, n, \size \in \nats \ldotp \forall
+        w_{\var{ret}} \in \Words \ldotp \\
+        \quad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
+        \quad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
+        \qquad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
+        \qquad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
+        \qquad \Rightarrow \\
+        \qquad \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
+        \qquad \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
+        \qquad \qquad \quad \Phi \step[j] \Phi' \land \\
+        \qquad \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
+        \qquad \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
+        \qquad \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
+        \qquad \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
+        \qquad \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+        \qquad \qquad \quad \size - 1 = e'-b'
+      \end{array}
+      \right) \land\\
+      \quad \left(
+      \begin{array}{l}
+        \forall \Phi \in \ExecConfs \ldotp\\
+        \quad(\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
+        \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
+        \qquad \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed
+      \end{array}
+        \right)
     \end{array}
   \]
-      \item \label{item:malloc-spec:fail} For all $\Phi \in \ExecConfs$,
-        \[
-          \begin{array}{l}
-            \text{If } (\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
-            \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
-            \text{Then } \exists j \in \nats \ldotp \Phi \step[j] \failed
-          \end{array}
-        \]
-      \end{enumerate}
-    \end{enumerate}
 \end{definition}
-
-  \caption{The specification of malloc.}
-  \label{fig:malloc-spec}
-\end{figure}
-As described previously, the specification of malloc ensures that malloc has no
+As described previously, the specification of malloc ensure that malloc has no
 capabilities pointing \emph{out} of malloc. It does, however, not say anything about
-capabilities that point \emph{in} to malloc. We obviously cannot
+capabilities that point \emph{in} to malloc. We obviously can obviously not
 allow this if we want to trust malloc. We have chosen to keep the malloc
@@
 Malloc should not just be available to trusted programs, but also to possibly
-malicious programs. This is safe as it follows from the specification that the malloc
-capability is always safe in a world with the malloc region:
+malicious programs. The specification should therefore entail that the malloc
+capability is safe which would make it safe to let even malicious programs have
+access to it.
 \begin{lemma}[Malloc is safe to pass to adversary]
   \label{lem:malloc-in-vr}
-  For all $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
+  For $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
@@
 
-The reason we allow trusted code and adversary to invoke malloc is just to make
-our work more realistic, but we are otherwise not interested in its details. As
-such, we do not give a malloc implementation. We are, however, confident that it
-is possible to make an implementation of malloc that satisfies the malloc
-specification in Definition~\ref{spec:malloc}. There are in fact two
-simplifications in our system that makes things easier: First, we do not consider
-deallocation of memory which means that the data structure malloc uses to keep
-track of free memory does not have to handle reclaimed memory. Second, the
-malloc specification does not permit malloc to run out of memory and thus refuse
-allocation. This is possible on our simple capability machine because it has an
-infinite address space. An initial capability with an infinite range of
-authority would of course need to be part of malloc, but it could also double as
-the data structure that keeps track of free memory. % \itoplassug{If we have
-  % references for malloc being difficult, we could add that to the text.}
-%\itoplassug{Add the assumption that the malloc capability in examples
-%  satisfies specification.}
-
-\section{Reusable macro instructions}
-\label{sec:macros}
+\itoplassug{Add the assumption that the malloc capability in examples
+  satisfies specification.}
+
+\subsection{Reusable macro instructions}
+\label{subsec:macros}
 %\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
@@
 I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
-With the calling convention and logical relation defined, we would like to show its usefulness by proving
-the correctness of a series of examples that rely
-on well-bracketedness and local-state encapsulation and use the calling
-convention to enforce these properties. However, the programs that
-run on our capability machine are assembly programs and even program examples that would
-be small in a high-level language become big and unintelligible at this
-low level. Thus to make our program examples intelligible, we introduce a
-series of low-level abstractions in the form of macros.
 We define a number of reusable macros capturing the calling convention as well
@@
 available in register $r_\stk$.
-\itoplassug{Lau, suggestion: should we mention temporary registers? They are
-  used in the macros, but it seems like a minor technical detail.}
 
 The macro \texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} captures
-those parts of the calling convention related to actually transferring control
-to some adversarial code. Specifically, it pushes the contents of the private
-registers, $\overline{r_{\var{priv}}}$, to the stack. Then it pushes the
-``restoration'' code to the stack. The restoration code will be executed as the
-first thing upon return, and restores the stack pointer and the old program
-counter. After the restoration code is pushed to the stack, a copy of the pc is
-pushed to the stack after it has been adjusted to point to the first instruction
-after the jump. From the stack pointer, a protected return pointer is created by
-adjusting it to point to the return code and encapsulating it by restricting its
-permission to $\entry$. Next, the stack pointer is restricted to the unused
-part, and the unused part of the stack is cleared (as discussed in
-Section~\ref{sec:stack-and-return-pointer}). Finally, the non-argument registers
-are cleared and we jump to $r$. Upon return after the restoration code has been
-executed, the restoration code is popped from the stack and the private words we
-pushed to the stack before the call are popped to the private registers. The
-implementation of \texttt{scall} is presented in
-\figurename~\ref{fig:scall-impl}, and the restoration code is presented in
-\figurename~\ref{fig:scall-rest-code}. The implementation of \texttt{scall} uses
-some of the macros we present next.
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} captures the
+parts of the calling convention related to actually transferring control to some
+adversarial code. Specifically, it pushes the contents of the private registers,
+$\overline{r_{\var{priv}}}$, to the stack. Then it pushes the restoration code
+to the stack. The restoration code will be executed as the first thing upon
+return, where it restores the stack pointer and the old program pointer. After
+the restoration code is pushed to the stack, a copy of the pc is adjusted to
+point to the first instruction after the control has been transferred and then
+pushed to the stack. From the stack pointer, a protected return pointer is
+created by adjusting it to point to the return code and encapsulating it by
+restricting its permission to enter. Next, the stack pointer is restricted to
+the unused part, and the unused part of the stack is cleared, so it does not
+govern any part of the private stack frame which means that it can safely be
+given away. Finally, the non-argument registers are cleared and we jump to $r$.
+Upon return after the restoration code has been executed, the restoration code
+is popped from the stack and the private words we pushed to the stack before the
+call are popped to the private registers. The implementation of \texttt{scall}
+is presented in \figurename~\ref{fig:scall-impl}, and the restoration code is
+presented in \figurename~\ref{fig:scall-rest-code}. The implementation of
+\texttt{scall} uses some of the macros we present later in this section.
+
+The macro \texttt{\footnotesize{mclear $r$}} clears all the memory the
+capability in register $r$ has authority over. It is used by \texttt{scall} to
+clear the unused part of the stack before control is transferred. It should also
+be used to clear the stack before returning to adversarial code. Similarly, the
+macro \texttt{\footnotesize{rclear $R$}} clears all the registers in the set
+$R$. It is also used by \texttt{scall}, and it should also be used before
+returning to adversarial code.
+
+The macros \texttt{\footnotesize{prepstack $r$}} and \texttt{\footnotesize{reqglob
+    $r$}} are the last macros related to the calling convention. The former,
+\texttt{prepstack} should be used when one receives a stack from an unknown
+sure. The \texttt{prepstack} macro first ensures the stack has permission
+read/write-local/execute, and then it adjusts the stack capability, so it
+follows the convention for the stack\footnote{The stack capability should always
+point to the top most word on the stack. A stack received from an unknown source
+can be treated as empty, so the stack capability should point just outside its
+range of authority.}. The other macro, \texttt{reqglob}, ensures that the
+capability in register $r$ is global. This macro should be used on callbacks
+received from unknown sources to ensure that they cannot be derived from the
+stack pointer.
+
+The remaining macros are not directly related to the calling convention, but
+they help making the examples in \sectionname~\ref{sec:examples} intelligible.
+The macros \texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}}
+add and remove elements from the stack. The macro \texttt{\footnotesize{fetch
+    $r$ $\var{name}$}} fetches the capability related to $\var{name}$ from the
+linking table and stores it in register $r$. The macro
+\texttt{\footnotesize{crtcls $\overline{(x_i,r_i)}$ $r$}} allocates a closure
+where $r$ points to the closure's code and a new environment is allocated (using
+malloc) where the contents of $\overline{r_i}$ is stored. In the code referred
+to by $r$, an implicit fetch happens when an instruction refers to $x_i$.
+
+The Appendix contains the implementation of all the macros used in
+\texttt{scall}. The technical appendix~\citep{technical_appendix} contains detailed descriptions
+of all the macros as well as all implementations.
+%\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
 
 \begin{figure}[htb]
-  \begin{toplas}
   \centering
@@
   move r_0 r_stk
-  lea r_0 $\var{ofc_\var{rec}}$ // -5 is the offset to the first instruction of
-the recovery code
+  lea r_0 $\var{ofc_\var{act}}$ // -5 is the offset to the first instruction of the activation code
   restrict r_0 $\encodePermPair((\local,\entry))$
@@
   \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
-      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
-      restoration code is given in \figurename~\ref{fig:scall-rest-code}. The
-      variable $\var{ofc}$ is the offset to the label after, and
-      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
-      the activation record. The set $R = \RegName \setminus
+    r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The restoration
+  code is given in \figurename~\ref{fig:scall-rest-code}. The variable
+  $\var{ofc}$ is the offset to the label after, and $\var{ofc_\var{act}} =
+  -5$ which is the offset to the first instruction of the activation record. The
+set $R = \RegName \setminus
 \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
   \label{fig:scall-impl}
-  \end{toplas}
-
 \end{figure}
 \begin{figure}[htb]
-  \begin{toplas}
   \centering
@@
 \end{lstlisting}
-    \caption{The restoration code used in \texttt{scall}. The variable
-      $\var{ofc}$ is 5 which is the offset to the address where the old stack
-      pointer is stored on the stack.}
+  \caption{The restoration code used in \texttt{scall}. The variable $\var{ofc}$ is 5
+    which is the offset to the address where the old stack pointer is stored on
+    the stack.} 
   \label{fig:scall-rest-code}
+\end{figure}
+
 \end{toplas}
-\end{figure}
-
-The macro \texttt{\footnotesize{mclear $r$}} clears all the memory the
-capability in register $r$ has authority over. It is used by \texttt{scall} to
-clear the unused part of the stack before control is transferred. It should also
-be used to clear the stack before returning to adversarial code. Similarly, the
-macro \texttt{\footnotesize{rclear $R$}} clears all the registers in the set
-$R$. It is also used by \texttt{scall}, and it should also be used before
-returning to adversarial code.
-
-The macros \texttt{\footnotesize{prepstack $r$}} and \texttt{\footnotesize{reqglob
-    $r$}} are the last macros related to the calling convention. The former,
-\texttt{prepstack} should be used when one receives a stack from an unknown
-source. The \texttt{prepstack} macro first ensures the stack has permission
-read/write-local/execute, and then it adjusts the stack capability, so it
-follows the convention for the stack\footnote{The stack capability should always
-point to the topmost word on the stack. A stack received from an unknown source
-can be treated as empty, so the stack capability should point just outside its
-range of authority.}. The other macro, \texttt{reqglob}, ensures that the
-capability in register $r$ is global. This macro should be used on callbacks
-received from unknown sources to ensure that they cannot be derived from the
-stack pointer.
-
-The remaining macros are not directly related to the calling convention, but
-they help making the examples in \sectionname~\ref{sec:examples} intelligible.
-The macros \texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}}
-add and remove elements from the stack. The macro \texttt{\footnotesize{fetch
-    $r$ $\var{name}$}} fetches the capability related to $\var{name}$ from the
-linking table and stores it in register $r$. The macro
-\texttt{\footnotesize{malloc $r$ $n$}} invokes malloc with size argument $n$.
-The \texttt{malloc} macro assumes that a capability for malloc resides in the
-linking table and is basically a fetch of the malloc capability followed by a
-setup of a return pointer. Finally, the macro \texttt{\footnotesize{crtcls
-    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
-closure's code and a new environment is allocated (using malloc) and the
-contents of $\overline{r_i}$ is stored in the environment. In the code referred
-to by $r$, an implicit load from the environment happens when an instruction
-refers to $x_i$.
-
-The Appendix contains the implementation of all the macros used in
-\texttt{scall}. The technical appendix~\citep{technical_appendix} contains more
-detailed descriptions of all the macros as well as all implementations.
-%\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
-We stress that the macros correspond to series of instructions as seen in
-\figurename~\ref{fig:scall-impl} and \ref{fig:scall-rest-code}, and that they
-are introduced for intelligibility. The examples of
-\sectionname~\ref{sec:examples}, the program examples are stated using the
-macros, but the proofs work on the expanded macros.
-\end{toplas}
 
 \begin{toplas}
-\section{Reasoning about programs on a capability machine}
+\section{Reasoning about programs}
 \label{sec:reasoning}
-% \itoplassug{
-%   \begin{itemize}
-%   \item A section with the lemmas we use for reasoning about programs
-%     \begin{itemize}
-%     \item lemma 58, scall works. Include description of how it handles the
-%       common things in call and leaves the argumentation for callee code and
-%       continuation.
-%     \item lemma 59, malloc works. If this is included, we should probably also
-%       include the malloc specification.
-%     \item lemma 60, create closure works.
-%     \item Include a description about how the FTLR is used to reason about
-%       unknown code.
-%     \item Mention that this is a step on the way towards having a program logic,
-%       and then in the discussion discuss why a program logic for this language
-%       is not what we need.
-%     \end{itemize}
-%   \end{itemize}
-% }
-There are many details to get right when programming in assembly. These details
-carry over to proofs about assembly programs, so many of the proofs in the next
-section about example programs are a bit cumbersome. It is especially annoying,
-when the same line of reasoning is applied in multiple places. To mitigate this,
-we have defined a number of lemmas that capture common reasoning patterns in
-these proofs. We present the most important such lemmas in this section, to give
-the reader an idea of how they work.
-
-% Anti-reduction lemma
-Our capability machine only allows the final memory of an execution to be
-observed, so naturally the correctness lemmas we prove in
-\sectionname~\ref{sec:examples} are statements about the memory in the halted
-configuration. In order to prove something about some part of the final memory,
-we establish an invariant for it as a region in a world and show that the
-initial configuration is in the $\observations$-relation w.r.t.\ that world. The
-$\observations$-relation says that if the initial configuration halts
-successfully, then the memory in the halting state must still respect a private
-future world of the initial world. It is, however, bothersome and error prone to
-try to argue about the entire execution in one go. Instead we want to reason
-modularly, in the sense that we only want to reason about parts of the execution
-at a time. To this end, we prove an anti-reduction lemma that essentially says
-that if we can show that an initial configuration, $\Phi$, steps to a
-configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then
-$\Phi$ is also in the observation relation.
-\begin{lemma}[Anti-reduction for $\observations$]
-\label{lem:anti-red-obs}
-  \begin{align*}
-    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
-    & \quad n' \geq n - i \land W' \futurestr W \land \\
-    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
-    & \quad \npair[n']{(\reg',\ms')} \in \observations(W') \\
-    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
-  \end{align*}
-\end{lemma}
-
-% FTLR
-The anti-reduction works well when we reason about the execution of known code,
-because we can execute it. When we want to reason about unknown code, the
-anti-reduction lemma does not apply because we do not know what instructions are
-being executed. This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
-As a reminder, the FTLR says that if a capability only has access to safe values
-with respect to a world, then it is safe to use it for execution in the same
-world. The unknown code we consider will be assumed to only have access to safe
-values, which typically means that we assume it has access to a linking table
-with a malloc capability and otherwise consists of instructions. This allows us
-to use the FTLR seeing as malloc is a safe value (cf.
-Lemma~\ref{lem:malloc-in-vr}) and instructions are really just integers and they
-are always safe values. Here, it is important to remember that the FTLR gives us
-that the capability for the untrusted code is safe w.r.t. some world, so even
-though we described the usual assumption on the unknown code's memory, it is the
-semantic model of the memory, i.e.\ the world, that the FTLR considers. In order
-to argue that a specific configuration is safe, we need to argue that the
-configuration is safe with respect to the world. That is, we need to argue that
-the memory satisfies the world and the register-file is in the
-$\stdrr$-relation. For instance, in the case where the untrusted code assumes
-control first, we will use the FTLR to show that the capability for the unknown
-code can be used for execution. The unknown code will get access to some known,
-trusted code through a capability in the initial register-file. Because the
-capability is in the register-file, we will have to argue that it is safe which
-we do using Lemma~\ref{lem:anti-red-obs}.
-
-% Many details to get right when reasoning about low-level programs.
-Another common pattern in proofs about programs on our capability machine
-concerns the use of scall. The following lemma captures the commonalities of
-reasoning about programs using scall. From another point of view, it can be seen
-as a specification for scall.
-\begin{lemma}[$\mathtt{scall}$ works]
-  \label{lem:scall-works}
-  If
-  \begin{enumerate}
-  \item \label{item:scall:memsat} $\memSat[n]{\ms}{\revokeTemp{W}}$ 
-  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
-  \item \label{item:scall:stack} $(\reg,\ms) \text{ is looking at }
-    \mathtt{scall}\;r(\overline{r_{\mathit{arg}}},
-    \overline{r_{\mathit{priv}}}) \text{ followed by }
-    c_{\mathit{next}}$\footnote{Defined in the appendix.}
-  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
-  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} If
+\itoplassug{
   \begin{itemize}
-    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{rec}}
-      \uplus \ms_{\mathit{unused}}')$,
-    \item $W' =
-      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
-    \item $\memSat[n-1]{\ms''}{W'}$
-    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
-    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
-      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r \mapsto \reg(r)]$ 
-    \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
-    \item $\npair[n-1]{c_\stk} \in \stdvr(W')$
+  \item A section with the lemmas we use for reasoning about programs
+    \begin{itemize}
+    \item lemma 58, scall works. Include description of how it handles the
+      common things in call and leaves the argumentation for callee code and
+      continuation.
+    \item lemma 59, malloc works. If this is included, we should probably also
+      include the malloc specification.
+    \item lemma 60, create closure works.
+    \item Include a description about how the FTLR is used to reason about
+      unknown code.
+    \item Mention that this is a step on the way towards having a program logic,
+      and then in the discussion discuss why a program logic for this language
+      is not what we need.
     \end{itemize}
-
-    then we have that $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
-  \item \label{item:scall:hyp-cont} \emph{Hyp-Cont} If
-    \begin{itemize}
-    \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}$
-    \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
-    \item for all $r$, we have that:
-      \begin{equation*}
-        \reg'(r)
-        \begin{cases}
-          = c_{\mathit{next}} &\text{ if } r = \pcreg\\
-          = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
-          \in \stdvr(\revokeTemp{W''}) &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\}
-        \end{cases}
-      \end{equation*}
-    \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
   \end{itemize}
-
-    then we have that $\npair[n']{(\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')} \in \observations(W'')$
-  \end{enumerate}
-  Then 
-  \begin{itemize}
-    \item $\npair{(\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}})} \in \observations(W)$
-  \end{itemize}
-
-  % Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
-  % words for stack pointer and return pointer and some fixed number of words for
-  % the standard return instructions).
-\end{lemma}
-Roughly, the \texttt{scall} lemma states that an invocation of scall is safe if
-\texttt{scall} is executed in a reasonable state
-(\ref{item:scall:memsat}-\ref{item:scall:stack}), the callee is safe to execute
-(\ref{item:scall:hyp-callee}), and returning to the code after the scall is safe
-(\ref{item:scall:hyp-cont}).
-
-In the common case \texttt{scall} is used to invoke untrusted code, and we will
-only have very general assumptions about that code, in which case we will use
-the FTLR (Theorem~\ref{thm:ftlr}) to argue
-Assumption~\ref{item:scall:hyp-callee}, using what we know about the values it
-has access to (e.g.\ linking table, malloc capability etc.). By assumption in
-Hyp-Callee, the memory is safe, so it suffices to show that the register-file
-contains safe values, which amounts to showing that the arguments in the call
-are safe.
-
-By using Lemma~\ref{lem:scall-works} to reason about \texttt{scall}, the proofs
-become agnostic to the actual implementation of \texttt{scall}. In other words,
-should we change the implementation of \texttt{scall}, then we just need to
-prove that Lemma~\ref{lem:scall-works} holds for the new implementation in order
-to gain that all our results still hold true.
-
-Also often used is the \texttt{malloc} instruction, so we also prove
-Lemma~\ref{lem:malloc-works} to help reason about it. The structure of the
-\texttt{malloc} lemma is close to that of the \texttt{scall} lemma. It also has
-requirements on the configuration just before \texttt{malloc} is executed
-(\ref{item:malloc:looking-at}-\ref{item:malloc:malloc-sat}) as well as
-requirements on the execution afterwards (\ref{item:malloc:hyp-cont}). It does
-not have any requirements on the callee as the \texttt{malloc} specification
-defines its behavior.
-\begin{lemma}[$\mathtt{malloc}$ works]
-  \label{lem:malloc-works}
-  If
-  \begin{enumerate}
-  \item \label{item:malloc:looking-at}\lookingat{(\reg,\ms)}{\mathtt{malloc}\;r\;k}{c_{\var{next}}}
-  \item $k \geq 0$
-  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
-  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
-  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
-  \item $\memSat{\ms}{W}$
-  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
-  \item \label{item:malloc:malloc-sat} $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
-  \item \label{item:malloc:hyp-cont} \emph{Hyp-Cont}
-    If
-    \begin{itemize}
-    \item $n' \leq n-1$
-    \item $\iota_\malloc \futurewk W(i)$
-    \item $\memSat[n']{\ms_{\var{footprint}}' \uplus \ms'}{W[i \mapsto \iota_\malloc]}$
-    \item $\memSat[n']{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
-      \begin{equation*}
-        \reg'(r') = 
-        \begin{cases}
-          c_{\var{next}} & r' = \pcreg \\
-          ((\rwx,\glob),\start,\addrend,\addr) & r' = r \\
-          \reg(r) & r' \not\in \RegName_t \union \{\pcreg, r, r_1\}
-        \end{cases}
-      \end{equation*}
-    \item $\addrend - \start = k - 1$
-    \item $\dom(\ms_{\var{alloc}}) = [\start,\addrend]$
-    \item $\forall \addr \in [\start,\addrend]\ldotp \ms_{\var{alloc}}(\addr) = 0$
-    \end{itemize}
-    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}}'  \uplus \ms_{\var{alloc}})} \in \observations(W[\iota_\malloc])$
-  \end{enumerate}
-  Then
-  \[
-    \npair{(\reg,\ms)} \in \observations(W)
-  \]
-\end{lemma}
-In the technical appendix~\citep{technical_appendix}, we also define a lemma for
-the macro used to create closures, \texttt{crtcls}.
-% The below remark seems unnecessary: isn't this true about any lemma in any proof?
-% We stress that these lemmas
-% are conveniences and are not strictly necessary to reason about programs for the
-% capability machine. In principle, one could reason solely in the semantic model
-% without the use of these lemmas, but that would lead to a lot of repetition.
+}
+
 \end{toplas}
@@
 lemma statements. The examples may look deceivingly short, but it is because
-they use the macro instructions described in \sectionname~\ref{sec:macros}.
+they use the macro instructions described in \sectionname~\ref{subsec:macros}.
 The examples would be
@@
 configuration is safe (in the $\observations$ relation) for a world with a
-permanent region that requires the assertion flag to be 0. By
-Lemma~\ref{lem:anti-red-obs}, it suffices to show that the configuration is safe
-after some reduction steps. We then use the \texttt{scall} lemma
-(Lemma~\ref{lem:scall-works}), by which it suffices to show that (1) the
+permanent region that requires the assertion flag to be 0. By an anti-reduction
+lemma, it suffices to show that the configuration is safe after some reduction
+steps. We then use a general lemma for reasoning about
+\texttt{\footnotesize{scall}}, by which it suffices to show that (1) the
 configuration that \texttt{\footnotesize{scall}} will jump to is safe and (2)
@@
 up is safe. We use the Fundamental Theorem to reason about the unknown
-adversarial code as described in \sectionname~\ref{sec:reasoning}, but notice
-that the adversary capability is an enter capability, which the Fundamental
-Theorem says nothing about. Luckily the enter capability becomes $\exec$ after
-the jump and then the Fundamental Theorem applies.
+adversarial code, but notice that the adversary capability is an enter
+capability, which the Fundamental Theorem says nothing about. Luckily the enter
+capability becomes $\exec$ after the jump and then the Fundamental Theorem
+applies.
 
@@
 
-\lau{Reviewer C, esop, would like a better comparison with \citet{Devriese:2016ObjCap} (but also admits that they did not follow the below discussion).}
+\lau{Reviewer C, esop, would like a better comparison with \citec{Devriese:2016ObjCap} (but also admits that they did not follow the below discussion).}
 Our logical relation is a unary, step-indexed Kripke logical relation with
@@
 \begin{toplas}
-In the future we would also like to investigate a new program logic for
-reasoning about capability safety for our capability machine model. In fact,
-we think the lemmas in Section~\ref{sec:reasoning} are suggestive of the style
-of results that could be written in such a logic. We think Iris would also be
-a natural starting point for such an endeavour, since Iris is really a
-framework, which can be instantiated to different programming languages. OCPL
-was able to leverage existing Iris specifications for the high-level language;
-for our capability machine model, however, it would be necessary to devise new
-kinds of specifications for our low-level programs with unstructured
-control-flow. It is likely that we could get inspiration from earlier work on
-logics for assembly programming languages, such as XCAP~\citep{xcap}.
+In the future we would also like to
+investigate a new program logic for reasoning about capability safety for our
+capability machine model. We think Iris would also be a natural starting point
+for such an endeavour, since Iris is really a framework, which can be
+instantiated to different programming languages. OCPL was able to leverage
+existing Iris specifications for the high-level language; for our capability
+machine model, however, it would be necessary to devise new kinds of
+specifications for our low-level programs with unstructured control-flow. It is
+likely that we could get inspiration from earlier work on logics for assembly
+programming languages, such as XCAP~\citep{xcap}.
 \end{toplas}
@@
 \section{Appendix}
-In this appendix, we give more precise formulations of lemmas that were mentioned in the paper, and the most important supporting definitions and lemmas.
-The goal is to provide details that can help to understand in more detail what we discuss in the paper. 
-Full details and proofs are not given here, but for those we refer to the technical appendix~\citep{technical_appendix}.
-
 \subsection{Logical relation}
@@
   \end{itemize}
+  We say that an o.f.e. $\left(X, \seqn{\nequal{n}}\right)$ is \emph{inhabited} if there
+  exists an element $x \in X$.
+\end{definition}
+
+If you are familiar with metric spaces observe that o.f.e.'s are but a different
+presentation of bisected $1$-bounded ultrametric spaces.
+
+\begin{definition}[Cauchy sequences and limits]
+  \label{def:cauchy-sequence}
+  Let $\left(X, \seqn{\nequal{n}}\right)$ be an o.f.e. and $\seq{x}$ be a sequence of
+  elements of $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
+  \begin{align*}
+    \forall k \in \NN, \exists j \in \NN, \forall n \geq j, x_j \nequal{k} x_n
+  \end{align*}
+  or in words, the elements of the chain get arbitrarily close.
+
+  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
+  \begin{align*}
+    \forall k \in \NN, \exists j \in \NN, \forall n \geq j, x \nequal{k} x_n.
+  \end{align*}
+  A sequence may or may not have a limit. If it has we say that the sequence
+  \emph{converges}. The limit is necessarily unique in this case
+  (Exercise~\ref{ex:limits-unique}) and we write $\lim_{n\to\infty}x_n$ for it.
 \end{definition}
@@
 
-\begin{definition}[Later on c.o.f.e.'s]
-  Let $\left( X, \left( \nequal \right)_{n=0}^{\infty}\right)$ be a \cofe{}. Define
-  \[
-  \blater \left( X, \left( \nequal \right)_{n=0}^{\infty}\right) = \left( X, \left(\nequal' \right)_{n=0}^{\infty}\right)
-  \] 
-  where for all $x$ and $x'$ in $X$
-  \[
-  x \nequal' x'\textit{  iff  } \begin{cases}
-    \top              & \text{if } n=0 \\
-    x \nequal[n-1] x' & \text{if } n > 0
-  \end{cases}
-  \]
-%  The set in the \cofe{} remains the same, but the equality relation is ``moved'' so it is equal at an earlier time.
-\todo[inline]{can we here say that to look under a later we have to use a computation step?}
+\begin{definition}
+  \label{def:nonexpansive-contractive-ofe}
+  Let $\left(X, \seqn{\nequal{n}_X}\right)$ and $\left(Y, \seqn{\nequal{n}_Y}\right)$ be
+  two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
+  The function $f$ is 
+  \begin{itemize}
+  \item \emph{non-expansive} if for any $x, x' \in X$, and any $n \in \NN$,
+    \begin{align*}
+      x \nequal{n}_X x' \implies f(x) \nequal{n}_Y f(x')
+    \end{align*}
+  \item \emph{contractive} if for any $x, x' \in X$, and any $n \in \NN$,
+    \begin{align*}
+      x \nequal{n}_X x' \implies f(x) \nequal{n+1}_Y f(x')
+    \end{align*}
+  \end{itemize}
 \end{definition}
+
+\begin{theorem}[Banach's fixed point theorem]
+  \label{thm:banach-fixed-point}
+  Let $\left(X, \seqn{\nequal{n}}\right)$ be a an inhabited c.o.f.e. and $f : X \to X$ a contractive
+  function. Then $f$ has a unique fixed point.
+\end{theorem}
+
+\begin{definition}[The category $\Ul$]
+  The category $\Ul$ of complete ordered families of equivalences has as objects complete
+  ordered families of equivalences and as morphisms non-expansive functions.
+\end{definition}
+
+From now on, we often use the underlying set $X$ to denote a (complete) o.f.e. $\left(X,
+  \seqn{\nequal{n}_X}\right)$, leaving the family of equivalence
+relations implicit.
+
+\begin{definition}
+  \label{def:later-functor}
+  The functor $\Later$ is a functor on $\Ul$ defined as
+  \begin{align*}
+    \Later\left(X, \seqn{{\nequal{n}}}\right) &= \left(X,
+          \seqn{\overset{n}{\equiv}}\right)\\
+    \Later(f) &= f
+  \end{align*}
+  where $\overset{0}{\equiv}$ is the total relation and $x \overset{n+1}{\equiv} x'$
+  iff $x \nequal{n} x'$
+\end{definition}
+
+\begin{definition}
+ \label{def:nonexpansive-loc-contractive-functor}
+ A functor $F : \op{\Ul} \times \Ul \to \Ul$ is \emph{locally non-expansive}  if for all
+ objects $X$, $X'$, $Y$, and $Y'$ in $\Ul$ and
+ $f, f' \in \hom{\Ul}{X}{X'}$ and $g, g' \in \hom{\Ul}{Y'}{Y}$ we have
+ \begin{align*}
+   f \overset{n}{=} f' \land g \overset{n}{=} g' \implies F(f, g) \overset{n}{=} F(f', g').
+ \end{align*}
+
+ It is \emph{locally contractive} if the stronger implication
+ \begin{align*}
+   f \overset{n}{=} f' \land g \overset{n}{=} g' \implies F(f, g) \overset{n+1}{=} F(f', g').
+ \end{align*}
+ holds. Note that the equalities are equalities on function spaces.
+\end{definition}
+
+\begin{proposition}
+  \label{prop:loc-nonexp-loc-contr}
+  If $F$ is a locally non-expansive functor then $\Later \comp F$ and $F \comp \left(\op{\Later}
+  \times \Later\right)$ are locally contractive. Here, the functor $F \comp \left(\op{\Later} \times
+  \Later\right)$ works as
+  \begin{align*}
+    (F \comp (\op{\Later} \times \Later))(X, Y) = F\left(\op{\Later}(X), \Later(Y)\right)
+  \end{align*}
+  on objects and analogously on morphisms and
+  $\op{\Later} : \op{\Ul} \to \op{\Ul}$ is just $\Later$ working on $\op{\Ul}$ (i.e., its
+  definition is the same).
+\end{proposition}
+
+\begin{definition}
+ \label{def:fixed-point}
+ A fixed point of a locally contractive functor $F$ is an object $X \in \Ul$, such that
+ $F(X, X) \iso X$.
+\end{definition}
+
+The following is America and Rutten's fixed point theorem~\cite{America-Rutten:JCSS89}.
+\begin{theorem}
+ \label{thm:contr-functors-have-fixed-points}
+ Every locally contractive functor $F$ such that $F(1, 1)$ is inhabited has a unique fixed
+ point. The fixed point is unique among inhabited c.o.f.e.'s.
+ If in addition $F(\emptyset, \emptyset)$ is inhabited then the fixed point of $F$ is unique.
+\end{theorem}
+In~\cite{BirkedalL:metric-enriched-journal} one can find a
+category-theoretic generalization, which shows how to obtain fixed
+points of locally contractive funtors on categories enriched in $\Ul$,
+in particular on the category of preordered c.o.f.e.'s.
 
@@
  \subsection{Example correctness lemmas}
-\begin{lemma}[Correctness lemma for \texttt{f1}, copy of Lemma~\ref{lem:correctness-f1}] \forcenewline
+\begin{lemma}[Correctness lemma for \texttt{f1}] \forcenewline
   \label{lem:correctness-f1-app}
@@
 
-\begin{lemma}[Correctness lemma for \texttt{f2}, detailed version of Lemma~\ref{lem:correctness-f2}]
+\begin{lemma}[Correctness lemma for \texttt{f2}]
   \label{lem:correctness-f2-app}
@@
 \end{lemma}
-\begin{lemma}[Correctness lemma for \texttt{f3}, detailed version of Lemma~\ref{lem:correctness-f3}]
-  \label{lem:correctness-f3-detailed}
+\begin{lemma}[Correctness lemma for \texttt{f3}]
+  \label{lem:correctness-f3}
   For all $n \in \nats$
@@
 
- \begin{lemma}[Correctness of $g1$, detailed version of Lemma~\ref{lem:correctness-g1}]
-  \label{lem:correctness-g1-detailed}
+ \begin{lemma}[Correctness of $g1$]
+  \label{lem:correctness-g1}
   For all $n \in \nats$
@@
 
-\subsection{Reasoning about programs definitions}
-\begin{definition}
-  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ 
-  iff
-  \begin{itemize}
-  \item $\reg(\pcreg) = ((p,g),b,e,a)$
-  \item $p = \rwx$, $p = \exec$, or $p = \rwlx$
-  \item $a+n\leq e$, $b\leq a\leq e$
-  \item $\ms(a+0,\cdots,a+n) = [i_0,\cdots,i_n]$
-  \item $c_{\mathit{next}} = ((p,g),b,e,a+n+1)$
-  \end{itemize}
-\end{definition}
-
-\begin{definition}
-  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c_\malloc}'' 
-  iff
-  \begin{itemize}
-  \item $\reg(pc) = \stdcap$
-  \item $\ms(\start) = ((\_,\_),\start_\link,\_,\_)$
-  \item $\ms(\start_\link+j) = c$
-  \end{itemize}
-\end{definition}
-
-\begin{definition}
-  We say that $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
-  iff
-  \begin{itemize}
-  \item $\reg(r_\stk) =((\rwlx,\local),b_\stk,e_\stk,a_\stk)$
-  \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
-  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$ \lau{Maybe make it clear what happens when $\ms_\stk$ is empty}
-  \item $b_\stk - 1\leq a_\stk$
-  \end{itemize}
-\end{definition}
-
 \end{document}

### 2018-05-08 toplas/paper.tex
added cofe stuff
Fixed compilation errors.

@@
 \subsection{Complete ordered family of equivalences (c.o.f.e)}
+\newcommand{\seq}[1]{\ensuremath{\left\{#1_n\right\}_{n=0}^{\infty}}}
+\newcommand{\seqn}[1]{\ensuremath{\left(#1\right)_{n=0}^{\infty}}}
+\newcommand{\NN}{\ensuremath{\mathbb{N}}}
+\newcommand{\Ul}{\ensuremath{\mathcal{U}}}
+\newcommand{\Later}{\ensuremath{\blacktriangleright}}
+\newcommand{\op}[1]{\ensuremath{#1^{\text{op}}}}
+\newcommand{\comp}{\circ}
+\newcommand{\iso}{\cong}
+\todo[inline]{Insert the missing references in the bibtex file.}
+
 \label{app:cofe}
@@
   \begin{align*}
-    \forall k \in \NN, \exists j \in \NN, \forall n \geq j, x_j \nequal{k} x_n
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal{k} x_n
   \end{align*}
@@
   \begin{align*}
-    \forall k \in \NN, \exists j \in \NN, \forall n \geq j, x \nequal{k} x_n.
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal{k} x_n.
   \end{align*}
@@
   \begin{itemize}
-  \item \emph{non-expansive} if for any $x, x' \in X$, and any $n \in \NN$,
+  \item \emph{non-expansive} if for any $x, x' \in X$, and any $n \in \nats$,
     \begin{align*}
@@
     \end{align*}
-  \item \emph{contractive} if for any $x, x' \in X$, and any $n \in \NN$,
+  \item \emph{contractive} if for any $x, x' \in X$, and any $n \in \nats$,
     \begin{align*}
@@
 
-The following is America and Rutten's fixed point theorem~\cite{America-Rutten:JCSS89}.
+The following is America and Rutten's fixed point theorem~\citep{America-Rutten:JCSS89}.
 \begin{theorem}
@@
 \end{theorem}
-In~\cite{BirkedalL:metric-enriched-journal} one can find a
+In~\citet{BirkedalL:metric-enriched-journal} one can find a
 category-theoretic generalization, which shows how to obtain fixed

### 2018-05-08 toplas/paper.tex
Inserted a missing reference.
Added the last missing reference
Removed reference to exercise
Reordered things in Appendix to follow structure of paper
Explain how the requirement for stack clearing applies to any revocation of local capabilities.

@@
 
+The need for stack clearing in our calling convention is an instance of a general caveat when using CHERI's local capabilities as a restricted form of capability revocation.
+Consider how our use of local capabilities can be interpreted as temporarily delegating the stack and return capability to callees and then revoking the granted authority after the callee returns.
+From this perspective, local capabilities are a general feature enabling this temporary delegation of authority for the duration of an invocation and this is also how they are described by the authors~\citep{Watson2015Cheri}.
+However, our requirement for stack clearing on boundary crossings is also general.
+Revoking authority that was granted temporarily using local capabilities, requires clearing all memory for which the invokee had write-local authority (or at least erasing all local capabilities from that memory).
+Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenario's where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
+In future work, we are planning to investigate an alternative notion of \emph{linear} capabilities that offer a limited form of revocation without such this limitation.
+
+Note, by the way, that CheriBSD's use of local capabilities in ccall does not actually involve a form of revocation.
+CheriBSD's model involves a trusted stack manager that gives every compartment access to its own private stack using a local, write-local capability \citep{Watson2015Cheri}.
+The locality of the stack capability allows the trusted stack manager to prevent compartments from leaking their stack pointer in a boundary crossing, but those capabilities are never actually revoked.
+In fact, a compartment can easily store away such local capabilities in its own private stack and recover them there during future invocations.
+
 \emph{Modularity} It is important that our calling convention is modular, i.e.\
@@
 the logical relation again tells us when this is safe. Namely, the logical
-relation says that we can only pass on safe arguments. For instance, when we
-receive a stack pointer from an adversary, then we may at some point want to
-pass on part of this stack pointer to, say, a callback. In order to do so, we
-need to make sure the stack pointer is safe which means that, if we have revoked
-temporary invariants, the stack must not directly or indirectly allow access to
-local values that we cannot guarantee safety of. When received from an
-adversary, we have to consider the contents of the stack unsafe, so before we
-pass it on, we have to clear it, or perform a dynamic safety analysis of the
-stack contents and anything it points to. Clearing everything is not always
-desirable and a dynamic safety analysis is hard to get right and potentially
-expensive.
+relation says that we can only pass on arguments that are safe in the world we
+invoke the adversary in. For instance, when we receive a stack pointer from an
+adversary, then we may at some point want to pass on part of this stack pointer
+to, say, a callback. In order to do so, we need to make sure the stack pointer
+is safe which means that, if we have revoked temporary invariants, the stack
+must not directly or indirectly allow access to local values that we cannot
+guarantee safety of. When received from an adversary, we have to consider the
+contents of the stack unsafe, so before we pass it on, we have to clear it, or
+perform a dynamic safety analysis of the stack contents and anything it points
+to. Clearing everything is not always desirable and a dynamic safety analysis is
+hard to get right and potentially expensive.
 
@@
 \newcommand{\iso}{\cong}
-\todo[inline]{Insert the missing references in the bibtex file.}
 
@@
   \emph{converges}. The limit is necessarily unique in this case
-  (Exercise~\ref{ex:limits-unique}) and we write $\lim_{n\to\infty}x_n$ for it.
+  and we write $\lim_{n\to\infty}x_n$ for it.
 \end{definition}
@@
 in particular on the category of preordered c.o.f.e.'s.
+
+\subsection{Load instruction sufficiency lemma}
+ \begin{lemma}[Conditions for store instruction are sufficient]
+   \label{lem:conds-store-suff}
+   If 
+   \begin{itemize}
+   \item $\ms = \ms' \uplus \ms_f$
+   \item $\heapSat[\ms']{n}{W}$
+   \item $((\perm,\gl),\start,\addrend,\addr) = c$
+   \item $\npair{c}\in\stdvr(W)$
+   \item $\writeAllowed{\perm}$
+   \item $\withinBounds{\var{c}}$
+   \item $\npair{\var{w}}\in\stdvr(W)$
+   \item if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in
+     \{\rwlx,\readwritel \}$
+   \end{itemize}
+
+   then $\addr \in \dom(\ms')$ (i.e. $\ms\update{a}{w} =
+   \ms'\update{a}{w}\uplus\ms_f$) and
+   $\heapSat[{\ms'\update{\addr}{\var{w}}}]{n}{W}$
+ \end{lemma}
 
@@
 respectively.
-
-\subsection{Sufficiency lemma}
- \begin{lemma}[Conditions for store instruction are sufficient]
-   \label{lem:conds-store-suff}
-   If 
-   \begin{itemize}
-   \item $\ms = \ms' \uplus \ms_f$
-   \item $\heapSat[\ms']{n}{W}$
-   \item $((\perm,\gl),\start,\addrend,\addr) = c$
-   \item $\npair{c}\in\stdvr(W)$
-   \item $\writeAllowed{\perm}$
-   \item $\withinBounds{\var{c}}$
-   \item $\npair{\var{w}}\in\stdvr(W)$
-   \item if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in
-     \{\rwlx,\readwritel \}$
-   \end{itemize}
-
-   then $\addr \in \dom(\ms')$ (i.e. $\ms\update{a}{w} =
-   \ms'\update{a}{w}\uplus\ms_f$) and
-   $\heapSat[{\ms'\update{\addr}{\var{w}}}]{n}{W}$
- \end{lemma}
 

### 2018-05-08 toplas/paper.tex
typos

@@
 Revoking authority that was granted temporarily using local capabilities, requires clearing all memory for which the invokee had write-local authority (or at least erasing all local capabilities from that memory).
-Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenario's where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
-In future work, we are planning to investigate an alternative notion of \emph{linear} capabilities that offer a limited form of revocation without such this limitation.
+Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenarios where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
+In future work, we are planning to investigate an alternative notion of \emph{linear} capabilities that offer a limited form of revocation without such limitations.
 

### 2018-05-08 toplas/paper.tex
Added first draft for presentation illustrations (caption and references missing!)

@@
 % \newtheorem{lemma}{Lemma}
+\input{preamble-illu}
 
@@
 
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}{0.32\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+
+    % Unused part
+      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,5) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,5)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+  \end{tikzpicture}
+  \caption{test}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=red] {\footnotesize ostk};
+      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=red] {\footnotesize opc};
+      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=red] {\footnotesize ...};
+      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=red] {\footnotesize ...};
+
+    % Unused part
+      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,2)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+  \end{tikzpicture}
+  \caption{test}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=teal] {\footnotesize 0};
+      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=teal] {\footnotesize 0};
+      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=teal] {\footnotesize 0};
+      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=teal] {\footnotesize 0};
+
+    % Unused part
+      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,2)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+  \end{tikzpicture}
+  \caption{test}
+  \end{subfigure}
+
+  \begin{subfigure}{0.32\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \inactsf{(0,5)}{(4.5,7)}{Our stack frame}
+
+    % second attacker
+      \actadv{(0,7)}{(4.5,10)}{Callee\\\footnotesize stack frame}
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,5) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,5)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+  \end{tikzpicture}
+  \caption{test}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5] {\footnotesize 0};
+      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5] {\footnotesize 0};
+      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5] {\footnotesize 0};
+      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5] {\footnotesize 0};
+
+    % Unknown part
+      \node[color=red] at (2.25,9) {\huge ?};
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,2)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+  \end{tikzpicture}
+  \caption{test}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+      \foreach \x in {5,5.5,...,10}
+      {
+        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,2)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+  \end{tikzpicture}
+  \caption{test}
+  \end{subfigure}
+
+  \caption{test}
+  \label{fig:ret-adv}
+\end{figure}
 \textbf{Invoking the adversary} A slightly more complex case is invoking the

### 2018-05-08 toplas/paper.tex
Added some captions and references for the new figures.
The captions need some work, but they illustrate the main idea.

@@
-\documentclass[format=acmsmall, review=false, screen=true]{acmart}
+\documentclass[format=acmsmall, review=true, screen=true]{acmart}
 
@@
 \textbf{Returning to the adversary} Perhaps the simplest situation is returning
-to the adversary after they invoked our code. In this case, we have received a
+to the adversary after they invoked our code (\figurename~\ref{fig:illu-intro}). In this case, we have received a
 return pointer from them, and we just need to jump to it as usual. An obvious
@@
 registers before we jump (since they may contain data or capabilities that the
-adversary should not get access to). Additionally, we may have used the stack
+adversary should not get access to, \figurename~\ref{fig:illu-priv-data}). Additionally, we may have used the stack
 for various purposes (register spilling, storing local state when invoking other
 functions etc.), so we also need to clear that data before returning to the
-adversary.
+adversary (\figurename~\ref{fig:illu-clear-priv}).
 
 However, if we are returning from a function that has itself invoked adversary
-code, then clearing the used part of the stack is not enough. The \emph{unused}
+code (\figurename~\ref{fig:illu-call-adv}), then clearing the used part of the stack is not enough. The \emph{unused}
 part of the stack may also contain data and capabilities, left there by the
@@
 back. In this case, the adversary could use the unused part of the stack to
-store local pointers and load them from there after they get control back. To
+store local pointers and load them from there after they get control back (\figurename~\ref{fig:illu-adv-data}). To
 prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
 the stack that the adversary has had access to, not just the parts that we have
-used ourselves. Since we may be talking about a large part of memory, this
+used ourselves (\figurename~\ref{fig:illu-clear-all}). Since we may be talking about a large part of memory, this
 requirement is the most problematic aspect of our calling convention for
@@
   \end{tikzpicture}
-  \caption{test}
+  \caption{We have been called by an adversary and are ready to return to the
+    adversary.}
+  \label{fig:illu-intro}
   \end{subfigure}
@@
   \end{tikzpicture}
-  \caption{test}
+  \caption{If we do not clear our stack when we return to an adversary, then
+    they can access our private data and capabilities. This could include old
+    stack and program pointers.}
+  \label{fig:illu-priv-data}
   \end{subfigure}
@@
   \end{tikzpicture}
-  \caption{test}
+  \caption{Here we made sure to clear the stack before returning to the
+    adversary. The adversary no longer has access to our private data and capabilities.}
+  \label{fig:illu-clear-priv}
   \end{subfigure}
@@
   \end{tikzpicture}
-  \caption{test}
+  \caption{In this case, we have called an adversary as part of our program.}
+  \label{fig:illu-call-adv}
   \end{subfigure}
@@
   \end{tikzpicture}
-  \caption{test}
+  \caption{If we return to the adversary after the situation in
+    \figurename~\ref{fig:illu-call-adv}, the adversary has access to the
+    contents of the stack from the call to the adversary. This could include the
+    return pointer, we gave to the adversary.}
+  \label{fig:illu-adv-data}
   \end{subfigure}
@@
   \end{tikzpicture}
-  \caption{test}
+  \caption{Here we made sure to clear all of the stack, including the part we
+    gave the adversary access to.}
+  \label{fig:illu-clear-all}
   \end{subfigure}
 
-  \caption{test}
+  \caption{In this section, we illustrate the stack in different scenarios
+    related to returning to an adversary. }
   \label{fig:ret-adv}

### 2018-05-08 toplas/preamble-illu.tex
preamble

--- 2017-12-18 presentations/illustrations/preamble.tex
@@
 \usepackage{tikz}
+\usepackage{tikzpeople}
 \usetikzlibrary{backgrounds}
+\usetikzlibrary{fit}
 \usetikzlibrary{calc}
@@
 \usetikzlibrary{arrows,shapes}
-\usepackage{tikzpeople}
 \usepackage{amssymb}
@@
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw ($#2 + (2.25,1.5)$) node[devil,mirrored,minimum size=1cm] {};
     \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5,align=center] {\footnotesize #4};
   }{
@@
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw ($#2 + (2.25,1.5)$) node[devil,mirrored,minimum size=1cm] {};
     \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
@@
   \scope
-    \clip (-.1,-.1) rectangle (6.1,15.1);
-    \fill[fill=white] (0,0) rectangle (6,15);
-    \draw (0,0) -- (0,15);
-    \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+    \draw[fill=gray!50] (0,-.5) rectangle (4.5,2)
+    node[pos=.5,color=black,align=center,yshift=0.1cm] {\footnotesize Higher
+      stack\\\footnotesize frames...};
   \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+  % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+  \draw (-1.5,0) node {};
+  \draw (-1.5,15) node {};
 }
@@
 \makeatother
+
+%%% Local Variables:
+%%% TeX-master: "paper"
+%%% End:

### 2018-05-09 toplas/paper.tex
Updated the illustration figure.
Removed individual captions and added a short
caption for all of the figure.

@@
         \clip (4.6,5) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,5)}{(4.6,13)}
+        \capbrace[sp]{(4.6,5)}{(4.6,12)}
       \end{scope}
@@
     \end{tikzpicture}
-  \caption{We have been called by an adversary and are ready to return to the
-    adversary.}
+    \caption{}
   \label{fig:illu-intro}
@@
       \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,2)}{(4.6,12)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+    \end{tikzpicture}
+  \caption{}
+  \label{fig:illu-priv-data}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=teal] {\footnotesize 0};
+      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=teal] {\footnotesize 0};
+      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=teal] {\footnotesize 0};
+      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=teal] {\footnotesize 0};
+
+    % Unused part
+      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,2)}{(4.6,12)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+    \end{tikzpicture}
+  \caption{}
+  \label{fig:illu-clear-priv}
+  \end{subfigure}
+
+  \begin{subfigure}{0.32\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \inactsf{(0,5)}{(4.5,7)}{Our stack frame}
+
+    % second attacker
+      \actadv{(0,7)}{(4.5,10)}{Callee\\\footnotesize stack frame}
+
+      % Stack pointer
+      \begin{scope}
+        \clip (4.6,5) rectangle (7.6,11.5);
+        \capbrace[sp]{(4.6,5)}{(4.6,13)}
+      \end{scope}
+      \draw node[right=0cm of sp] { sp };
+    \end{tikzpicture}
+  \caption{}
+  \label{fig:illu-call-adv}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % trusted stack frame
+      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
+      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5] {\footnotesize 0};
+      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5] {\footnotesize 0};
+      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5] {\footnotesize 0};
+      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5] {\footnotesize 0};
+
+    % Unknown part
+      \node[color=red] at (2.25,9) {\huge ?};
 
@@
     \end{tikzpicture}
-  \caption{If we do not clear our stack when we return to an adversary, then
-    they can access our private data and capabilities. This could include old
-    stack and program pointers.}
-  \label{fig:illu-priv-data}
+  \caption{}
+  \label{fig:illu-adv-data}
   \end{subfigure}
@@
       \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=teal] {\footnotesize 0};
-      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=teal] {\footnotesize 0};
-      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=teal] {\footnotesize 0};
-      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=teal] {\footnotesize 0};
-
-    % Unused part
-      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
+      \foreach \x in {5,5.5,...,10}
+      {
+        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
+      };
 
@@
     \end{tikzpicture}
-  \caption{Here we made sure to clear the stack before returning to the
-    adversary. The adversary no longer has access to our private data and capabilities.}
-  \label{fig:illu-clear-priv}
-  \end{subfigure}
-
-  \begin{subfigure}{0.32\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \inactsf{(0,5)}{(4.5,7)}{Our stack frame}
-
-    % second attacker
-      \actadv{(0,7)}{(4.5,10)}{Callee\\\footnotesize stack frame}
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,5) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,5)}{(4.6,13)}
-      \end{scope}
-      \draw node[right=0cm of sp] { sp };
-  \end{tikzpicture}
-  \caption{In this case, we have called an adversary as part of our program.}
-  \label{fig:illu-call-adv}
-  \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5] {\footnotesize 0};
-      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5] {\footnotesize 0};
-      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5] {\footnotesize 0};
-      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5] {\footnotesize 0};
-
-    % Unknown part
-      \node[color=red] at (2.25,9) {\huge ?};
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,2)}{(4.6,13)}
-      \end{scope}
-      \draw node[right=0cm of sp] { sp };
-  \end{tikzpicture}
-  \caption{If we return to the adversary after the situation in
-    \figurename~\ref{fig:illu-call-adv}, the adversary has access to the
-    contents of the stack from the call to the adversary. This could include the
-    return pointer, we gave to the adversary.}
-  \label{fig:illu-adv-data}
-  \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \foreach \x in {5,5.5,...,10}
-      {
-        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,2)}{(4.6,13)}
-      \end{scope}
-      \draw node[right=0cm of sp] { sp };
-  \end{tikzpicture}
-  \caption{Here we made sure to clear all of the stack, including the part we
-    gave the adversary access to.}
+  \caption{}
   \label{fig:illu-clear-all}
   \end{subfigure}
-
-  \caption{In this section, we illustrate the stack in different scenarios
-    related to returning to an adversary. }
+  \caption{Illustration of the situations related to stack clearing when returning to an adversary.
+    The greyed out areas are parts of the stack that are not accessible at the time. }
   \label{fig:ret-adv}

### 2018-05-09 toplas/paper.tex
Added a figure that illustrates the last example for when the adv hides a stack pointer on the stack.

@@
         \clip (4.6,5) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,5)}{(4.6,12)}
+        \capbrace{(4.6,5)}{(4.6,12)}{sp}
       \end{scope}
-      \draw node[right=0cm of sp] { sp };
     \end{tikzpicture}
@@
         \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,2)}{(4.6,12)}
+        \capbrace{(4.6,2)}{(4.6,12)}{sp}
       \end{scope}
-      \draw node[right=0cm of sp] { sp };
     \end{tikzpicture}
@@
         \clip (4.6,5) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,5)}{(4.6,13)}
+        \capbrace{(4.6,5)}{(4.6,13)}{sp}
       \end{scope}
-      \draw node[right=0cm of sp] { sp };
     \end{tikzpicture}
@@
         \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace[sp]{(4.6,2)}{(4.6,13)}
+        \capbrace{(4.6,2)}{(4.6,13)}{sp}
       \end{scope}
-      \draw node[right=0cm of sp] { sp };
     \end{tikzpicture}
@@
   \caption{Illustration of the situations related to stack clearing when returning to an adversary.
-    The greyed out areas are parts of the stack that are not accessible at the time. }
+    The greyed out areas are parts of the stack that are not accessible at the time.}
   \label{fig:ret-adv}
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine the adversary stores their stack pointer, invokes trusted code
-that uses part of the stack to store private data and then invokes the adversary
+pointers. Imagine the adversary stores their stack pointer (\ref{fig:illu-adv-store-sp}), invokes trusted code
+that uses part of the stack to store private data (\ref{fig:illu-adv-call}) and then invokes the adversary
 again with a stack pointer restricted to exclude the part containing the private
-data. If the adversary had a way of keeping hold of their old stack pointer, it
+data (\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
 could access the private data stored there by the trusted code and break
@@
 % Domi: this is now discussed in the discussion section, no need to discuss it here?
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}{0.32\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
+    % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbracebot{(4.6,2)}{(4.6,12)}{sp}
+      \end{scope}
+    \end{tikzpicture}
+    \caption{}
+  \label{fig:illu-adv-store-sp}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
+
+    % Our stack frame
+      \actsf{(0,5)}{(4.5,6.5)}{Our stack frame}
+
+    % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbracebotred{(4.6,2)}{(4.6,12)}{sp}
+      \end{scope}
+      \begin{scope}
+        \clip (5.6,5) rectangle (9,11.5);
+        \capbrace{(5.6,5)}{(5.6,12)}{sp'}
+      \end{scope}
+
+    \end{tikzpicture}
+  \caption{}
+  \label{fig:illu-adv-call}
+  \end{subfigure}
+  \begin{subfigure}{0.333\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker ..
+      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
+
+    % Stored stack pointer
+      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
+
+    % Our stack frame
+      \inactsf{(0,5)}{(4.5,6.5)}{Our stack frame}
+
+    % Second attacker
+      \actadv{(0,6.5)}{(4.5,9.5)}{Callee \\ \footnotesize stack frame}
+
+    % Stack pointer
+      \begin{scope}
+        \clip (4.6,2) rectangle (7.6,11.5);
+        \capbracebotred{(4.6,2)}{(4.6,12)}{sp}
+      \end{scope}
+      \begin{scope}
+        \clip (5.6,6.6) rectangle (9,11.5);
+        \capbrace{(5.6,6.5)}{(5.6,12)}{sp'}
+      \end{scope}
+    \end{tikzpicture}
+  \caption{}
+  \label{fig:illu-adv-called}
+  \end{subfigure}
+  \caption{Illustration of the situations related to stack clearing when invoking an adversary.
+    The greyed out areas are parts of the stack that are not accessible at the time.
+  }
+  \label{fig:ret-adv}
+\end{figure}
 

### 2018-05-09 toplas/preamble-illu.tex
Added a figure that illustrates the last example for when the adv hides a stack pointer on the stack.

@@
 
-\newcommand{\capbrace}[3][sp1]{
+\newcommand{\capbrace}[4][sp1]{
   \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
+  #2 -- #3 node (#1) [black,midway,xshift=0.8cm] {\footnotesize #4};
+}
 
+\newcommand{\capbracebot}[4][sp1]{
+  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  #2 -- #3 node (#1) [black,near start,xshift=0.8cm] {\footnotesize #4};
+}
+
+\newcommand{\capbracebotred}[4][sp1]{
+  \draw [decorate,red,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  #2 -- #3 node (#1) [red,near start,xshift=0.8cm] {\footnotesize #4};
 }

### 2018-05-10 toplas/paper.tex
Added Rels definition to appendix.
Move a figure reference to a more appropriate place.

@@
 registers before we jump (since they may contain data or capabilities that the
-adversary should not get access to, \figurename~\ref{fig:illu-priv-data}). Additionally, we may have used the stack
+adversary should not get access to). Additionally, we may have used the stack
 for various purposes (register spilling, storing local state when invoking other
 functions etc.), so we also need to clear that data before returning to the
-adversary (\figurename~\ref{fig:illu-clear-priv}).
+adversary (\figurename~\ref{fig:illu-priv-data} and \figurename~\ref{fig:illu-clear-priv}).
 
@@
 \end{mathpar}
+%
+\textbf{Transition system relations}
+\begin{align*}
+   \Rels &= \{(\phi_\pub, \phi) \in \powerset{\States^2}\times \powerset{\States^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
+\end{align*}
+%
 \textbf{Erasure}
@@
 \end{align*}
+%
 \textbf{Active region projection}
@@
 \end{align*}
+%
 \textbf{Projection of regions based on locality}
@@
 \]
+%
 \textbf{Address stratification}

### 2018-05-10 toplas/paper.tex
A draft of the discussion about the stack clearing upon return.

@@
 \sectionname~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
-calling convention without security guarantees. However, most of our security
-measures require only a few atomic checks or register clearings on boundary
-crossings between trusted code and adversary, which should produce an acceptable
-performance overhead. The only exception are the requirements for stack clearing
-that we have in two situations: when returning to the adversary and when
-invoking an adversary callback. As we have explained, we need to clear all of
-the stack that we are not using ourselves, not just the part that we have
-actually used. In other words, on every boundary cross between trusted code and
-adversary code, a potentially large region of memory must be cleared. We believe
-this is actually a common requirement for typical usage scenarios of local
-capabilities and capability machines like CHERI should consider to provide
-special support for this requirement, in the form of a highly-optimized
-instruction for erasing a large block of memory. Nevertheless, from a discussion
-with the designers of the CHERI capability machine, we gather that it is not immediately
-clear whether and how such a primitive could be implemented efficiently in the CHERI context.
-\itoplassug{Maybe refer to some of the papers reviewer A, esop mentions for stack clearing. See comment in tex file. }
-% http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf
-% https://www.usenix.org/legacy/event/osdi08/tech/full_papers/zeldovich/zeldovich_html/index.html
-% http://users.ece.utexas.edu/~tiwari/pubs/MICRO-08-rangecache.pdf
-
-The need for stack clearing in our calling convention is an instance of a general caveat when using CHERI's local capabilities as a restricted form of capability revocation.
+calling convention without security guarantees. However, most require only a few
+atomic checks or register clearings on boundary crossings between trusted code
+and adversary, which should produce an acceptable performance overhead. The only
+exception are the requirements for stack clearing that we have in two
+situations: when returning to the adversary and when invoking an adversary
+callback. As we have explained, we need to clear all of the stack that we are
+not using ourselves, not just the part that we have actually used. In other
+words, on every boundary cross between trusted code and adversary code, a
+potentially large region of memory must be cleared.
+
+First, contrary to what we explained before, we actually suspect that this overhead can be avoided in one of the two cases: when returning to the adversary.
+In that case, we now think it would suffice to clear only the (much smaller) part of the stack used by the trusted code itself.
+To understand this, it is useful to take another look at the illustrations in Figure~\ref{fig:ret-adv} related to this case.
+If we do not clear the stack upon return (as illustrated in Figure~\ref{fig:illu-clear-all}), then the adversary might have used that stack to store local capabilities they received in a previous invocation (see Figure~\ref{fig:illu-adv-data}).
+In other words, the stack clearing is necessary for revoking those local capabilities: the stack pointer and return pointer for the invocation illustrated in Figure~\ref{fig:illu-adv-data}.
+While this approach is safe, we now suspect that we could do without the revocation in this case.
+The stack pointer which the adversary was given access to in the stack frame depicted in Figure~\ref{fig:illu-call-adv} only carries authority that the adversary has access to in the higher stack frame anyway.
+Similarly, the return pointer is merely an entry pointer pointing into the trusted code's stack frame, and this is also a capability that the adversary we return to could have constructed themselves from their stack pointer.
+
+This improvement of our calling convention is a recent insight and not yet reflected in our examples and our proofs.
+We do believe the proofs could be updated to accomodate for this change, but it would require some non-trivial changes.
+Consider the awkward example proof, for example, we would have to ensure that the world $W_6$, depicted in Figure~\ref{fig:worlds-in-pf-g1} would not just be a private future world of $W_5$, but a public one (in addition to being a public future world of $W_1$).
+This would allow us to do without the stack clearing, but it would entail some changes to the islands we use, and an extra argument that the old adversary's stack and return pointer remain valid after clearing the trusted code's stack frame and relaxing the invariant that used to ensure it could not be modified.
+While this change is a clear improvement, we do not actually believe that it fundamentally changes the efficiency characteristics of the approach: the cost is halved, but remains asymptotically the same.
+
+A second important remark we want to make is that the need for stack clearing in our calling convention is an instance of a general caveat when using CHERI's local capabilities as a restricted form of capability revocation.
 Consider how our use of local capabilities can be interpreted as temporarily delegating the stack and return capability to callees and then revoking the granted authority after the callee returns.
@@
 In fact, a compartment can easily store away such local capabilities in its own private stack and recover them there during future invocations.
+
+Since local capabilities seem intended to provide a restricted form of revocation, perhaps capability machines like CHERI should consider to provide special support for this requirement.
+Ideally, such support would take the form of a highly-optimized instruction for erasing a large block of memory.
+However, from a discussion with the designers of the CHERI capability machine, we gather that it is not immediately clear whether and how such a primitive could be implemented efficiently in the CHERI context.
+\itoplassug{Maybe refer to some of the papers
+  reviewer A, esop mentions for stack clearing. See comment in tex file. }
+% http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf
+% https://www.usenix.org/legacy/event/osdi08/tech/full_papers/zeldovich/zeldovich_html/index.html
+% http://users.ece.utexas.edu/~tiwari/pubs/MICRO-08-rangecache.pdf
 

### 2018-05-11 toplas/paper.tex
appendix, cofes

@@
 \newcommand{\iso}{\cong}
+\renewcommand{\hom}[3]{#1(#2,#3)}
 
@@
   \label{def:cauchy-sequence}
-  Let $\left(X, \seqn{\nequal{n}}\right)$ be an o.f.e. and $\seq{x}$ be a sequence of
+  Let $\left(X, \seqn{\nequal} \right)$ be an o.f.e. and $\seq{x}$ be a sequence of
   elements of $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
@@
   \label{def:nonexpansive-contractive-ofe}
-  Let $\left(X, \seqn{\nequal{n}_X}\right)$ and $\left(Y, \seqn{\nequal{n}_Y}\right)$ be
+  Let $\left(X, \seqn{\nequal_{X}}\right)$ and $\left(Y, \seqn{\nequal_{Y}}\right)$ be
   two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
@@
     \begin{align*}
-      x \nequal{n}_X x' \implies f(x) \nequal{n}_Y f(x')
+      x \nequal_X x' \implies f(x) \nequal_Y f(x')
     \end{align*}
@@
     \begin{align*}
-      x \nequal{n}_X x' \implies f(x) \nequal{n+1}_Y f(x')
+      x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x')
     \end{align*}
@@
   \label{thm:banach-fixed-point}
-  Let $\left(X, \seqn{\nequal{n}}\right)$ be a an inhabited c.o.f.e. and $f : X \to X$ a contractive
+  Let $\left(X, \seqn{\nequal}\right)$ be a an inhabited c.o.f.e. and $f : X \to X$ a contractive
   function. Then $f$ has a unique fixed point.
@@
 
-From now on, we often use the underlying set $X$ to denote a (complete) o.f.e. $\left(X,
-  \seqn{\nequal{n}_X}\right)$, leaving the family of equivalence
-relations implicit.
-
 \begin{definition}
@@
   \begin{align*}
-    \Later\left(X, \seqn{{\nequal{n}}}\right) &= \left(X,
+    \Later\left(X, \seqn{\nequal}\right) &= \left(X,
           \seqn{\overset{n}{\equiv}}\right)\\
@@
   where $\overset{0}{\equiv}$ is the total relation and $x \overset{n+1}{\equiv} x'$
-  iff $x \nequal{n} x'$
+  iff $x \nequal x'$
 \end{definition}
+
+From now on, we often use the underlying set $X$ to denote a (complete) o.f.e. $\left(X,
+  \seqn{\nequal}\right)$, leaving the family of equivalence
+relations implicit.
 
@@
 points of locally contractive funtors on categories enriched in $\Ul$,
-in particular on the category of preordered c.o.f.e.'s.
+in particular on the category of preordered c.o.f.e.'s.  A preodered c.o.f.e. is
+a c.o.f.e. equipped with a preorder that is closed under taking limits of
+converging sequences.
+The formulation in \emph{loc. cit.} also applies to solve
+mutually recursive domain equations on preordered c.o.f.e.'s; see~\citet{bizjak:mutually-recursive-mcat}
+for an explicit statement. That is the solution theorem we use to prove
+Theorem~\ref{thm:world-existence}.
 

### 2018-05-11 toplas/paper.tex
Minor things (label and removed empty tikz picture)
Moved things around in Appendix and added definitions related to the awkward example.

--- 2018-05-10 toplas/paper.tex
@@
 \newenvironment{toplas}
-    {\color{OliveGreen}
-
-    }{}
+    {\color{OliveGreen}}{}
 \newenvironment{toplassug}
-    {\color{Blue}
-
-    }{}
+    {\color{Blue}}{}
 
@@
     (1l) edge (2l);
-  \end{tikzpicture}
-  \begin{tikzpicture}[main node/.style={}]
   \end{tikzpicture}
@@
   }
-  \label{fig:ret-adv}
+  \label{fig:invoke-adv}
 \end{figure}
@@
 respectively.
+
+\subsection{Reasoning about programs definitions}
+\begin{definition}
+  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ 
+  iff
+  \begin{itemize}
+  \item $\reg(\pcreg) = ((p,g),b,e,a)$
+  \item $p = \rwx$, $p = \exec$, or $p = \rwlx$
+  \item $a+n\leq e$, $b\leq a\leq e$
+  \item $\ms(a+0,\cdots,a+n) = [i_0,\cdots,i_n]$
+  \item $c_{\mathit{next}} = ((p,g),b,e,a+n+1)$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}
+  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c_\malloc}'' 
+  iff
+  \begin{itemize}
+  \item $\reg(pc) = \stdcap$
+  \item $\ms(\start) = ((\_,\_),\start_\link,\_,\_)$
+  \item $\ms(\start_\link+j) = c$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}
+  We say that $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
+  iff
+  \begin{itemize}
+  \item $\reg(r_\stk) =((\rwlx,\local),b_\stk,e_\stk,a_\stk)$
+  \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
+  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$ \lau{Maybe make it clear what happens when $\ms_\stk$ is empty}
+  \item $b_\stk - 1\leq a_\stk$
+  \end{itemize}
+\end{definition}
 
@@
 
-\subsection{Reasoning about programs definitions}
+\subsection{Awkward example}
+\textbf{The region for $x$} The region $\iota_x$, is the region omitted from the proof sketch for the awkward example.
+\figurename~\ref{fig:x-reg} illustrates the transition system of $\iota_x$.
+  \begin{figure}[tb]
+    \centering
+\begin{tikzpicture}[->,shorten >=1pt,auto,node distance=3cm,
+  thick,main node/.style={circle,fill=white,draw,font=\sffamily\large},rect node/.style={rectangle,fill=blue!20,draw,font=\sffamily\large}]
+
+  \node[main node] (n0) {\footnotesize $\texttt{x}=0$};
+  \node (s) [left of=n0,xshift=1.75cm] {  };
+  \node[main node] (n1) [right of=n0]  {\footnotesize $\texttt{x}=1$};
+  \draw (s) edge[>=stealth] (n0);
+  \draw (n0) edge[bend left,>=stealth] (n1);
+  \draw (n1) edge[bend left,dashed,>=stealth] (n0);
+
+\end{tikzpicture}
+  \caption{Illustration of transition system in $\iota_x$. The dashed line is
+    the private transition.}
+  \label{fig:x-reg}
+\end{figure}
+
 \begin{definition}
-  We say that $(\reg,\ms) \text{ is looking at } [i_0,\cdots,i_n] \text{ followed by } c_{\mathit{next}}$ 
-  iff
-  \begin{itemize}
-  \item $\reg(\pcreg) = ((p,g),b,e,a)$
-  \item $p = \rwx$, $p = \exec$, or $p = \rwlx$
-  \item $a+n\leq e$, $b\leq a\leq e$
-  \item $\ms(a+0,\cdots,a+n) = [i_0,\cdots,i_n]$
-  \item $c_{\mathit{next}} = ((p,g),b,e,a+n+1)$
-  \end{itemize}
+  \label{def:iotax-region}
+  \begin{align*}
+    \iota_x   & = (\perm, 0, \phi_\pub, \phi, H_x) \\
+    \phi_\pub & = \{(0,1)\}^* \\
+    \phi      & = (1,0) \union \phi_\pub \\
+    H_x \; s \; \hat{W} & = \{\npair{\ms} \mid \ms(x) = s \land n > 0 \} \union \{\npair[0]{\ms}\}
+  \end{align*}
 \end{definition}
-
-\begin{definition}
-  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c_\malloc}'' 
-  iff
-  \begin{itemize}
-  \item $\reg(pc) = \stdcap$
-  \item $\ms(\start) = ((\_,\_),\start_\link,\_,\_)$
-  \item $\ms(\start_\link+j) = c$
-  \end{itemize}
-\end{definition}
-
-\begin{definition}
-  We say that $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
-  iff
-  \begin{itemize}
-  \item $\reg(r_\stk) =((\rwlx,\local),b_\stk,e_\stk,a_\stk)$
-  \item $\dom(\ms_{\mathit{unused}}) = [a_\stk+1,\cdots,e_\stk]$
-  \item $\dom(\ms_\stk) = [b_\stk,\cdots,a_\stk]$ \lau{Maybe make it clear what happens when $\ms_\stk$ is empty}
-  \item $b_\stk - 1\leq a_\stk$
-  \end{itemize}
-\end{definition}
+\textbf{Static region} This static region only requires that the memory segment is the given one.
+As it does not require safety, capabilities for this region cannot be gives to adversarial code.
+\begin{align*}
+  \iota^\sta (v,\ms) &= (v,1,=,=,H^\sta\;\ms)\\
+  H^\sta \; \ms \; s \; \hat{W} = & \{\npair{\ms} \mid n > 0 \} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
+\end{align*}
+\textbf{Static safe region} Static region that also requires safety.
+It is safe to give adversarial code read capabilities for this region.
+\begin{align*}
+  \iota^{\sta,u} (v,\ms) &= (v,1,=,=,H^{\sta,u}\;\ms)\\
+  H^{\sta,u} \; \ms \; s \; \hat{W} = & \left\{\npair{\ms'} \middle|
+    \begin{aligned}
+      &\ms' = \ms \land\\
+      &\forall \addr \in \dom(\ms) \ldotp\\
+      & \quad \nonlocal{\ms(\addr)} \land\\
+      & \quad \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))
+    \end{aligned}
+        \right\} \union \{\npair[0]{\ms'} \mid \ms' \in \Mems \}
+\end{align*}
 

### 2018-05-11 toplas/paper.tex
Very first draft for awkward example proof sketch

@@
 \end{lemma}
-
+\begin{toplas}
+  \begin{proof}[Proof sketch]
+    Define a world $W_1$ with the following regions
+    \begin{itemize}
+    \item A malloc region, $\iota_{\malloc,0}$.
+    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
+    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
+    \item A adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
+    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
+      accepts $\ms_\flag$ and $\ms_{g1}$.
+    \end{itemize}
+    If we can show
+    \begin{equation}
+      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
+  \end{equation}
+  \footnote{We will ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$.
+  In $W$ the flag region is permanent, so it guarantees that it will remain unchanged.
+  As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code pointed to by $c_\adv$.
+  With the standard no-write-local region chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
+  In order to get \ref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
+  We have defined the world, so there (almost) is a one-to-one correspondence between region and memory segment in the initial configuration, so (a) is easily shown.
+  In order to show (b), we need to show $c_\stk \in \stdvr(W)$ and $c_{g1} \in \stdvr(W)$.
+  the former follows easily from the choice of a permit-write-local region in $W$ for the stack.
+  In order to argue the latter, we basically have to argue that $c_{g1}$, our closure generator, respects the world $W$.
+  This amounts to showing that the closures generated by \texttt{g1} also respects the invariants.
+  The reasoning concerning variable $x$ follows that of \citet{Dreyer:jfp12}, so we will ignore it in the remainder of the proof and focus on the reasoning related to this low-level scenario.
+  The capability for the closure is a global enter capability that we call $c_{f4}$.
+  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary.
+  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg_1$ and memory $\ms_1$ satisfies a world $W_1$ that is a private future world of $W$ with regions prompted the closure generation.
+  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg_1[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms_1) \in \observations(W_1)$.
+  We know the first of the execution, so we use the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) to reason about the execution up until the first call.
+  During the first bit of execution, it is checked that the stack pointer has permission read/write-local/execute and that the callback is global.
+  Based on the stack pointers permission, we know that the capability is local which means that the region that governs it must be temporary.
+  This allows us to construct $W_2 \futurestr W_1$ where the stack region has been revoked (notice that it is also important for the reasoning about $x$ that this is a private future world).
+  In $W_2$, we replace the stack region with a temporary static region that governs our stack frame including the private registers we push to the stack and the stack recovery code as well as a standard region $\iote^{pwl}$-region for the remainder of the stack.
+  As the callback is global, it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
+  Notice that had the callback been local, this would not be the case.
+  We also need to reason about the return pointer we provide in the invocation.
+  The return pointer is a local capability, so we may assume that it is invoked in some $W_3 \futurewk W_2$.
+  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still there in $W_3$.
+  For the sake of presentation, we went through the above reasoning, but we actually us Lemma~\ref{lem:scall-works} to reason about \texttt{scall}.
+  To reason about the next bit of execution we know, we once again use Lemma~\ref{lem:anti-red-obs} followed by \ref{lem:scall-works} to handle \texttt{scall}.
+  As the private stack has changed, we need to replace the temporary static region that governs it, so it is revoked.
+  This means that the \texttt{scall} takes place in a world $W_4 \futurestr W_3$.
+  The reasoning about \texttt{scall} is the same as for the first callback invocation, so again we will get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
+  At this point, we can use Lemma~\ref{lem:anti-red-obs} to get to the point where we jump to the return pointer.
+  If we can argue that it is safe to jump to the return pointer we got initially from the adversary, then the proof is done.
+  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to otherwise the execution fails which is considered safe.}.
+  W.l.o.g.\ assume the return pointer is local.
+  This means that we have to use it in a world $W_6$ that is a public future world of $W_1$, which means that we have to restore all the regions in the world that the adversary relies on.
+  In particular, we will have to restore the stack region from $W_1$.
+  We cannot just pick $W_1$ because Lemma~\ref{lem:anti-red-obs} requires $W_6 \futurestr W_5$.
+  Before we proceed, let us consider where each of the worlds came from so far: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we chose $W_2$ and $W_4$ which means that we know exactly what changes they made.
+  In particular, we did not mask any of the revoked or temporary regions in $W_1$ with a permanent region.
+  This means that in $W_6$, we can reinstate every temporary and revoked region in $W_1$.
+  In the private future worlds we construct, we only take transitions that relative to $W_1$ are public.
+  The worlds we are given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is not problem with respect to the public future world relation.
+  In $W_5$ there will be regions to handle the stack.
+  These regions obviously need to be revoked as the $W_1$ stack regions replace them.
+  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world permits as it is extensional.
+  All in all, we create a world in which the invariants that the adversary relies on has been restored which makes it safe to invoke the return pointer.
+  \end{proof}
+\end{toplas}
 As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the

### 2018-05-11 toplas/paper.tex
removed obsolete sections

@@
 \end{toplas}
-As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
-macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
-is global, essentially to make sure it is not allocated on the stack where it
-might contain old stack pointers or return pointers. Otherwise, the
-encapsulation of our local stack frame could be broken. In the proof of
-Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
-callback in a world that is only a private future world of the one where we
-received the callback, precisely because we have invalidated the adversary's
-local state (particularly their old stack and return capabilities). The callback
-is still valid in this private future world, but only because we know that it is
-global.
-
-\itoplassug{Suggestion: Add an actual proof sketch}
-In Lemma~\ref{lem:correctness-g1} the order of control has been
-inverted compared to the previous lemmas. In this lemma, the adversary
-assumes control first with a capability for the closure creator
-\texttt{\footnotesize{g1}}. Consequently, we need to check that all
-arguments are safe to use and that we clean up before returning in the
-end. The inversion of control poses an interesting challenge when it
-comes to reasoning about the adversary's local state during the
-execution of \texttt{\footnotesize{f4}} and the callbacks where the
-adversary should not rely on the local state from before the call of
-\texttt{\footnotesize{f4}}. This is easily done by revoking all the
-temporary regions of the world given at the start of
-\texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}}
-returns, the adversary is again allowed to rely on its old local state
-so we need to guarantee that the local state is unchanged. This is
-important because the return pointer that \texttt{\footnotesize{f4}}
-receives may be local, and the adversary is allowed
-to allocate the activation record on the stack (just like we do) so
-they can store and recover their old stack pointer after
-\texttt{\footnotesize{f4}} returns.
-
-\begin{toplas}
-In \figurename~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
-the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
-for all the situations where control is passed to the adversary. The potentially
-local return pointer received by \texttt{\footnotesize{f4}} from the adversary
-is safe in $W_1$, so it can only be used in public future worlds. Let us see how
-$W_6$ is constructed as a public future world of $W_1$, as well as a private
-future world of $W_5$. The worlds $W_2$ and $W_4$ are constructed by revoking
-all temporary regions and adding a $\iota^\pwl$ region for the stack we pass
-away and static temporary regions (a region that accepts exactly one memory) for
-our local stack as well as the adversary's local stack. The given worlds $W_3$
-and $W_5$ are public future worlds in which the adversary chooses to return to
-us. None of these worlds can have masked the temporary regions of $W_1$ with
-permanent regions. $W_6$ is constructed by revoking the temporary regions of
-$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$ a public
-future world of $W_1$. It is therefore safe to use the return pointer to return
-to the adversary, since we have restored validity of any local state they might
-have stashed away.
+
+  The complete proof can be found in the technical appendix~\citep{technical_appendix}.
 \end{toplas}
+% As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
+% macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
+% is global, essentially to make sure it is not allocated on the stack where it
+% might contain old stack pointers or return pointers. Otherwise, the
+% encapsulation of our local stack frame could be broken. In the proof of
+% Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
+% callback in a world that is only a private future world of the one where we
+% received the callback, precisely because we have invalidated the adversary's
+% local state (particularly their old stack and return capabilities). The callback
+% is still valid in this private future world, but only because we know that it is
+% global.
+
+% \itoplassug{Suggestion: Add an actual proof sketch}
+% In Lemma~\ref{lem:correctness-g1} the order of control has been
+% inverted compared to the previous lemmas. In this lemma, the adversary
+% assumes control first with a capability for the closure creator
+% \texttt{\footnotesize{g1}}. Consequently, we need to check that all
+% arguments are safe to use and that we clean up before returning in the
+% end. The inversion of control poses an interesting challenge when it
+% comes to reasoning about the adversary's local state during the
+% execution of \texttt{\footnotesize{f4}} and the callbacks where the
+% adversary should not rely on the local state from before the call of
+% \texttt{\footnotesize{f4}}. This is easily done by revoking all the
+% temporary regions of the world given at the start of
+% \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}}
+% returns, the adversary is again allowed to rely on its old local state
+% so we need to guarantee that the local state is unchanged. This is
+% important because the return pointer that \texttt{\footnotesize{f4}}
+% receives may be local, and the adversary is allowed
+% to allocate the activation record on the stack (just like we do) so
+% they can store and recover their old stack pointer after
+% \texttt{\footnotesize{f4}} returns.
+
+% \begin{toplas}
+% In \figurename~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
+% the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
+% for all the situations where control is passed to the adversary. The potentially
+% local return pointer received by \texttt{\footnotesize{f4}} from the adversary
+% is safe in $W_1$, so it can only be used in public future worlds. Let us see how
+% $W_6$ is constructed as a public future world of $W_1$, as well as a private
+% future world of $W_5$. The worlds $W_2$ and $W_4$ are constructed by revoking
+% all temporary regions and adding a $\iota^\pwl$ region for the stack we pass
+% away and static temporary regions (a region that accepts exactly one memory) for
+% our local stack as well as the adversary's local stack. The given worlds $W_3$
+% and $W_5$ are public future worlds in which the adversary chooses to return to
+% us. None of these worlds can have masked the temporary regions of $W_1$ with
+% permanent regions. $W_6$ is constructed by revoking the temporary regions of
+% $W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$ a public
+% future world of $W_1$. It is therefore safe to use the return pointer to return
+% to the adversary, since we have restored validity of any local state they might
+% have stashed away.
+% \end{toplas}
 

### 2018-05-11 toplas/paper.tex
Updated awk. example proof sketch (possible compile error fixed)

@@
   \end{equation}
-  \footnote{We will ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$.
-  In $W$ the flag region is permanent, so it guarantees that it will remain unchanged.
+  \footnote{We will ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
   As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code pointed to by $c_\adv$.
@@
   This allows us to construct $W_2 \futurestr W_1$ where the stack region has been revoked (notice that it is also important for the reasoning about $x$ that this is a private future world).
-  In $W_2$, we replace the stack region with a temporary static region that governs our stack frame including the private registers we push to the stack and the stack recovery code as well as a standard region $\iote^{pwl}$-region for the remainder of the stack.
+  In $W_2$, we replace the stack region with a temporary static region that governs our stack frame including the private registers we push to the stack and the stack recovery code as well as a standard region $\iota^{pwl}$-region for the remainder of the stack.
   As the callback is global, it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
@@
   All in all, we create a world in which the invariants that the adversary relies on has been restored which makes it safe to invoke the return pointer.
+
+  The complete proof can be found in the technical appendix~\citep{technical_appendix}.
   \end{proof}
-\end{toplas}
-
-  The complete proof can be found in the technical appendix~\citep{technical_appendix}.
 \end{toplas}

### 2018-05-11 toplas/paper.tex
Some updates to the awk example proof sketch

@@
   \end{equation}
-  \footnote{We will ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
-  As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code pointed to by $c_\adv$.
-  With the standard no-write-local region chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
+  \footnote{We ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
+
+  As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code, so we use it to reason about $c_\adv$.
+  With the standard region $\iota^{nwl,p}$ chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
   In order to get \ref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
@@
   In order to show (b), we need to show $c_\stk \in \stdvr(W)$ and $c_{g1} \in \stdvr(W)$.
-  the former follows easily from the choice of a permit-write-local region in $W$ for the stack.
-  In order to argue the latter, we basically have to argue that $c_{g1}$, our closure generator, respects the world $W$.
-  This amounts to showing that the closures generated by \texttt{g1} also respects the invariants.
+  The former follows easily from the choice of a $\iota^{pwl}$-region for the stack in $W$.
+  In order to argue the latter, we basically have to argue that $c_{g1}$, the closure generator, respects the world $W$.
+  This amounts to showing that the closures generated by \texttt{g1} also respect the invariants of $W$.
   The reasoning concerning variable $x$ follows that of \citet{Dreyer:jfp12}, so we will ignore it in the remainder of the proof and focus on the reasoning related to this low-level scenario.
+
   The capability for the closure is a global enter capability that we call $c_{f4}$.
-  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary.
-  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg_1$ and memory $\ms_1$ satisfies a world $W_1$ that is a private future world of $W$ with regions prompted the closure generation.
-  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg_1[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms_1) \in \observations(W_1)$.
-  We know the first of the execution, so we use the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) to reason about the execution up until the first call.
+  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary, i.e.
+  $c_{f4} \in \stdvr(W')$ where $W'$ is $W$ with the regions added after executing \texttt{g1}.
+  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg$ and memory $\ms$ satisfies a world $W_1$ that is a private future world of $W'$, i.e.\ $\reg \in \stdrr(W_1)$ and $\memSat[]{\ms}{W_1}$.
+  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms) \in \observations(W_1)$.
+  When $c_{f4}$ is invoked, we know exactly what instructions are executed up until the \texttt{scall}, so we can use the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) to reason about this part.
   During the first bit of execution, it is checked that the stack pointer has permission read/write-local/execute and that the callback is global.
@@
 
+  For the sake of presentation, we have omitted many details and made several simplifications in the above proof.
   The complete proof can be found in the technical appendix~\citep{technical_appendix}.

### 2018-05-11 toplas/paper.tex
Finished pass over proof sketch.

@@
   We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms) \in \observations(W_1)$.
-  When $c_{f4}$ is invoked, we know exactly what instructions are executed up until the \texttt{scall}, so we can use the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}) to reason about this part.
-  During the first bit of execution, it is checked that the stack pointer has permission read/write-local/execute and that the callback is global.
-  Based on the stack pointers permission, we know that the capability is local which means that the region that governs it must be temporary.
-  This allows us to construct $W_2 \futurestr W_1$ where the stack region has been revoked (notice that it is also important for the reasoning about $x$ that this is a private future world).
-  In $W_2$, we replace the stack region with a temporary static region that governs our stack frame including the private registers we push to the stack and the stack recovery code as well as a standard region $\iota^{pwl}$-region for the remainder of the stack.
-  As the callback is global, it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
-  Notice that had the callback been local, this would not be the case.
-  We also need to reason about the return pointer we provide in the invocation.
-  The return pointer is a local capability, so we may assume that it is invoked in some $W_3 \futurewk W_2$.
-  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still there in $W_3$.
-  For the sake of presentation, we went through the above reasoning, but we actually us Lemma~\ref{lem:scall-works} to reason about \texttt{scall}.
-  To reason about the next bit of execution we know, we once again use Lemma~\ref{lem:anti-red-obs} followed by \ref{lem:scall-works} to handle \texttt{scall}.
-  As the private stack has changed, we need to replace the temporary static region that governs it, so it is revoked.
-  This means that the \texttt{scall} takes place in a world $W_4 \futurestr W_3$.
-  The reasoning about \texttt{scall} is the same as for the first callback invocation, so again we will get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
-  At this point, we can use Lemma~\ref{lem:anti-red-obs} to get to the point where we jump to the return pointer.
-  If we can argue that it is safe to jump to the return pointer we got initially from the adversary, then the proof is done.
+  When $c_{f4}$ is invoked, we know exactly what instructions are executed up until the \texttt{scall}, namely it is ensured that the callback is global and the stack pointer has read/write-local/execute permission, and $x$ is set to 0.
+  Because we know part of the execution, we can apply the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}).
+
+  The next part in the execution is an \texttt{scall}, so according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
+  For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
+  Based on the stack pointer's permission, we know that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{c.f.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
+  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.
+  In $W_2$, we replace the stack region of $W_1$ with a temporary static region that governs our stack frame, that is the private registers we push to the stack and the stack recovery code, as well as a standard region $\iota^{pwl}$-region for the part of the stack, we will provide to the callback.
+  The static region ensures that our local stack frame remains unchanged during the execution of the callback, and the standard region makes sure that the callback only puts safe values on the stack.
+  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
+  Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
+  This part corresponds to arguing about Hyp-Callee in the \texttt{scall} lemma.
+
+  We need to argue that it is safe to give the return pointer that \texttt{scall} constructs to the adversary which correspond to reason about the remainder of the execution of \texttt{f4} (corresponding to the Hyp-Cont case of the \texttt{scall} lemma).
+  The return pointer is a local capability, so we may assume that it is invoked in some configuration that satisfies $W_3 \futurewk W_2$.
+  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still present in $W_3$.
+  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 0), so for the part of the execution up until the \texttt{scall} we once again use the anti-reduction lemma.
+
+  For the \texttt{scall} itself we of course apply the \texttt{scall} lemma.
+  As the private stack has changed, we need to replace the two regions that govern the stack.
+  This means that the second invocation of the callback takes place in a world $W_4 \futurestr W_3$.
+  The reasoning about \texttt{scall} is the same as for the first callback invocation.
+  The callback capability is still safe because it is global which basically covers the Hyp-Callee case.
+  For the Hyp-Cont case, we get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
+  At this point, we can use the anti-reduction lemma a final time to get to the point where we jump to the return pointer.
+
+  If we can argue that it is safe to jump to the return pointer that we got initially from the adversary, then the proof is done.
   We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to otherwise the execution fails which is considered safe.}.
   W.l.o.g.\ assume the return pointer is local.
-  This means that we have to use it in a world $W_6$ that is a public future world of $W_1$, which means that we have to restore all the regions in the world that the adversary relies on.
-  In particular, we will have to restore the stack region from $W_1$.
-  We cannot just pick $W_1$ because Lemma~\ref{lem:anti-red-obs} requires $W_6 \futurestr W_5$.
-  Before we proceed, let us consider where each of the worlds came from so far: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we chose $W_2$ and $W_4$ which means that we know exactly what changes they made.
-  In particular, we did not mask any of the revoked or temporary regions in $W_1$ with a permanent region.
-  This means that in $W_6$, we can reinstate every temporary and revoked region in $W_1$.
-  In the private future worlds we construct, we only take transitions that relative to $W_1$ are public.
-  The worlds we are given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is not problem with respect to the public future world relation.
-  In $W_5$ there will be regions to handle the stack.
+  This means that it is safe in public future worlds of $W_1$ (remember that it came from $\reg$ and $\reg \in \stdrr(W_1)$).
+  This means that we need to construct a world $W_6$ with all the temporary and revoked regions of $W_1$ (this corresponds to restoring the invariants the adversary relies on for its safe execution).
+  Further for all these regions, $W_6$ must use the same region names as $W_1$ for $W_6$ to be a public future world.
+  As we used the anti-reduction lemma to get to this point, $W_6$ must also be constructed to be a private future world of $W_5$.
+  Before we construct $W_6$, let us consider where each of the worlds we have seen so far came from: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we constructed $W_2$ and $W_4$ which are the only private future worlds we have considered so far.
+  This means that we know exactly what changes they made.
+  In particular, we did not mask any of the temporary or revoked regions in $W_1$ with a permanent region.
+  This means that in $W_6$, we can reinstate every temporary or revoked region of $W_1$.
+  In the private future worlds we constructed, we only took transitions that, relative to $W_1$, are public.
+  The worlds we were given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is no problem with respect to the public future world relation.
+  In $W_5$ there were regions to handle the stack.
   These regions obviously need to be revoked as the $W_1$ stack regions replace them.
-  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world permits as it is extensional.
-  All in all, we create a world in which the invariants that the adversary relies on has been restored which makes it safe to invoke the return pointer.
+  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world relation permits as it is extensional.
+  All in all, it is possible to construct $W_6$, so it is both a public future world of $W_1$ and a private world of $W_5$ which means that it is indeed safe to return to the adversary.
 
@@
 that registers a redraw callback. The redraw callback gets temporary
-read-write access to a framebuffer.
+read/write access to a framebuffer.
 

### 2018-05-11 toplas/paper.tex
Added a paragraph outlining the new additions in this paper.

@@
 Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}
+
+\begin{toplas}
+  This paper is an extension of a published conference
+  paper\citep[Skorstengaard:esop2018]. We have made minor improvements to
+  readability and completeness throughout the paper, but we highlight the
+  following changes:
+  \begin{itemize}
+  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provide informal intuition about how the logical relation machinery comes into play in this setting.
+  \item We have added a proof sketch for the \nameref{thm:ftlr} and one for Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
+  \item We have added a section on malloc.
+    In particular, we provide the specification for malloc used in the examples of the paper.
+  \item We have added a section on macro instructions.
+    All the macro instructions are described and the implementation of \texttt{scall} is provided.
+  \item We have added a section on reasoning about programs that run on a capability machine.
+  \item We have added many details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
+    Many of the details we did not include in the paper are now in the appendix of the paper.
+  \end{itemize}
+\end{toplas}
 

### 2018-05-11 toplas/paper.tex
Added lemma about rwlx implies local capability

@@
 \end{lemma}
+
+\begin{toplas}
+  In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
+  This would indicate that our logical relation should imply the same, namely that capabilities with write-local permission are local:
+  \begin{lemma}[Stack capabilities are local]~
+    \label{lem:wl-local}
+    \begin{itemize}
+    \item If $\memSat{\ms}{W}$ and $\npair{\stdcap} \in \stdvr(W)$ and $\start
+      \leq \addrend$ and $\perm \in \{\rwlx,\rwl\}$, then $\gl = \local$.
+    \end{itemize}
+  \end{lemma}
+  In our definition of worlds, nothing prevents a world from having regions that are overlapping.
+  This is an issue for the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
+  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness as it prevents the regions from overlapping.
+  The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
+\end{toplas}
 

### 2018-05-11 toplas/paper.tex
Added reference to rwlx perm implies local in awk example proof sketch.

@@
   For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
-  Based on the stack pointer's permission, we know that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{c.f.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
+  Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{c.f.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
   This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.

### 2018-05-11 toplas/paper.tex
Added a footnote referring to the appendix for more on cofes and later.

@@
 the (recursive) equations in the theorem (where the $\blater$ operator can be
-safely ignored).
+safely ignored\footnote{The interested reader can find a brief coverage of \cofe{}'s
+and $\blater$ in the appendix.}).
 % Short technical explanation

### 2018-05-11 toplas/paper.tex
added figures as a point to introduction w/ new things
Changed all inline code to have normal text size instead of \footnotesize

@@
   \item We have added a proof sketch for the \nameref{thm:ftlr} and one for Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
+  \item We have added figures that illustrate central parts of the calling convention.
   \item We have added a section on malloc.
@@
 
-The macro \texttt{\footnotesize{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)}} captures
+The macro \texttt{scall
+    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures
 those parts of the calling convention related to actually transferring control
@@
 
-The macro \texttt{\footnotesize{mclear $r$}} clears all the memory the
+The macro \texttt{mclear $r$} clears all the memory the
 capability in register $r$ has authority over. It is used by \texttt{scall} to
@@
 be used to clear the stack before returning to adversarial code. Similarly, the
-macro \texttt{\footnotesize{rclear $R$}} clears all the registers in the set
+macro \texttt{rclear $R$} clears all the registers in the set
 $R$. It is also used by \texttt{scall}, and it should also be used before
@@
 
-The macros \texttt{\footnotesize{prepstack $r$}} and \texttt{\footnotesize{reqglob
-    $r$}} are the last macros related to the calling convention. The former,
+The macros \texttt{prepstack $r$} and \texttt{reqglob
+    $r$} are the last macros related to the calling convention. The former,
 \texttt{prepstack} should be used when one receives a stack from an unknown
@@
 they help making the examples in \sectionname~\ref{sec:examples} intelligible.
-The macros \texttt{\footnotesize{push $r$}} and \texttt{\footnotesize{pop $r$}}
-add and remove elements from the stack. The macro \texttt{\footnotesize{fetch
-    $r$ $\var{name}$}} fetches the capability related to $\var{name}$ from the
+The macros \texttt{push $r$} and \texttt{pop $r$}
+add and remove elements from the stack. The macro \texttt{fetch
+    $r$ $\var{name}$} fetches the capability related to $\var{name}$ from the
 linking table and stores it in register $r$. The macro
-\texttt{\footnotesize{malloc $r$ $n$}} invokes malloc with size argument $n$.
+\texttt{malloc $r$ $n$} invokes malloc with size argument $n$.
 The \texttt{malloc} macro assumes that a capability for malloc resides in the
 linking table and is basically a fetch of the malloc capability followed by a
-setup of a return pointer. Finally, the macro \texttt{\footnotesize{crtcls
-    $\overline{(x_i,r_i)}$ $r$}} allocates a closure where $r$ points to the
+setup of a return pointer. Finally, the macro \texttt{crtcls
+    $\overline{(x_i,r_i)}$ $r$} allocates a closure where $r$ points to the
 closure's code and a new environment is allocated (using malloc) and the
@@
 \subsection{Encapsulation of Local State}
-\texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
+\texttt{f1} and \texttt{f2} in
 \figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability
@@
 then test whether the local state is unchanged. They differ in the way
-they do this. Program \texttt{\footnotesize{f1}} uses our stack-based
-calling convention (captured by \texttt{\footnotesize{scall}}) to call
+they do this. Program \texttt{f1} uses our stack-based
+calling convention (captured by \texttt{scall}) to call
 the adversary, so it can use the available stack to store its local
-state.  On the other hand, \texttt{\footnotesize{f2}} uses malloc to
+state.  On the other hand, \texttt{f2} uses malloc to
 allocate memory for its local state and uses a calling convention based on heap
@@
 
-In order to formulate results about \texttt{\footnotesize{f1}} and
-\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
+In order to formulate results about \texttt{f1} and
+\texttt{f2}, we need a way to observe whether the assertion
 fails. To this end, we assume they have access to a flag (an address in memory).
 If the assertion fails, then the flag is set to $1$ and execution halts. The correctness lemma
-for \texttt{\footnotesize{f1}} then states:
+for \texttt{f1} then states:
 \begin{lemma}
@@
   \item $\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}$
-  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
+  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
   \item $\ms_\flag(\flag) = 0$
@@
 (Lemma~\ref{lem:scall-works}), by which it suffices to show that (1) the
-configuration that \texttt{\footnotesize{scall}} will jump to is safe and (2)
-that the configuration just after \texttt{\footnotesize{scall}} is done cleaning
+configuration that \texttt{scall} will jump to is safe and (2)
+that the configuration just after \texttt{scall} is done cleaning
 up is safe. We use the Fundamental Theorem to reason about the unknown
@@
 
-We have a similar lemma for \texttt{\footnotesize{f2}}:
+We have a similar lemma for \texttt{f2}:
 \begin{lemma}
@@
 \subsection{Well-Bracketed Control-Flow} 
-Using the stack-based calling convention of \texttt{\footnotesize{scall}}, we get
+Using the stack-based calling convention of \texttt{scall}, we get
 well-bracketed control-flow. To illustrate this, we look at
-two example programs \texttt{\footnotesize{f3}} and
-\texttt{\footnotesize{g1}} in \figurename~\ref{fig:prog-f3-and-g1}.
-
-In \texttt{\footnotesize{f3}} there are two calls to an adversary and in order
+two example programs \texttt{f3} and
+\texttt{g1} in \figurename~\ref{fig:prog-f3-and-g1}.
+
+In \texttt{f3} there are two calls to an adversary and in order
 for the assertion in the middle to succeed, they need to be well-bracketed. If
 the adversary were able to store the return pointer from the first call and
-invoke it in the second call, then \texttt{\footnotesize{f3}} would have $2$ on
+invoke it in the second call, then \texttt{f3} would have $2$ on
 top of its stack and the assertion would fail. However, the security measures in
@@
 
-The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}},
+The final example, \texttt{g1} with \texttt{f4},
 is a faithful translation of a tricky example known from the literature (known
 as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists
-of two parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
-\texttt{\footnotesize{g1}} is a closure generator that generates closures with
-one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
+of two parts, \texttt{g1} and \texttt{f4}.
+\texttt{g1} is a closure generator that generates closures with
+one variable $x$ set to $0$ in its environment and \texttt{f4} as
 the program (note we can omit some calling convention security measures because
-the stack is not used in the closure generator). \texttt{\footnotesize{f4}}
+the stack is not used in the closure generator). \texttt{f4}
 expects one argument, a callback. It sets $x$ to $0$ and calls the callback.
@@
 Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties,
-like the well-bracketedness that Example~\texttt{\footnotesize{g1}} relies on.
+like the well-bracketedness that Example~\texttt{g1} relies on.
 In fact, our examples show that our logical relation imply a stronger form of

### 2018-05-11 toplas/paper.tex
minor edits and comments on proof of awkward example

@@
   With the standard region $\iota^{nwl,p}$ chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
-  In order to get \ref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
+  In order to get \eqref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
   We have defined the world, so there (almost) is a one-to-one correspondence between region and memory segment in the initial configuration, so (a) is easily shown.
@@
   The reasoning concerning variable $x$ follows that of \citet{Dreyer:jfp12}, so we will ignore it in the remainder of the proof and focus on the reasoning related to this low-level scenario.
+\lars{preceding sentence is unclear: what is $x$, what does ``this low-level scenario'' refer to ?}
 
@@
 
-  We need to argue that it is safe to give the return pointer that \texttt{scall} constructs to the adversary which correspond to reason about the remainder of the execution of \texttt{f4} (corresponding to the Hyp-Cont case of the \texttt{scall} lemma).
+  We need to argue that it is safe to give the return pointer that \texttt{scall} constructs to the adversary which corresponds to reasoning about the remainder of the execution of \texttt{f4} (corresponding to the Hyp-Cont case of the \texttt{scall} lemma).
   The return pointer is a local capability, so we may assume that it is invoked in some configuration that satisfies $W_3 \futurewk W_2$.
@@
 
-  For the \texttt{scall} itself we of course apply the \texttt{scall} lemma.
+  For the \texttt{scall} itself we, of course, apply the \texttt{scall} lemma.
   As the private stack has changed, we need to replace the two regions that govern the stack.
@@
   If we can argue that it is safe to jump to the return pointer that we got initially from the adversary, then the proof is done.
-  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to otherwise the execution fails which is considered safe.}.
+  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to since otherwise the execution fails which is considered safe.}.
   W.l.o.g.\ assume the return pointer is local.
   This means that it is safe in public future worlds of $W_1$ (remember that it came from $\reg$ and $\reg \in \stdrr(W_1)$).
-  This means that we need to construct a world $W_6$ with all the temporary and revoked regions of $W_1$ (this corresponds to restoring the invariants the adversary relies on for its safe execution).
-  Further for all these regions, $W_6$ must use the same region names as $W_1$ for $W_6$ to be a public future world.
+  Hence we need to construct a world $W_6$ with all the temporary and revoked regions of $W_1$ (this corresponds to restoring the invariants the adversary relies on for its safe execution).
+  Further, for $W_6$ to be a public future world, for all these regions, $W_6$ must use the same region names as $W_1$. 
   As we used the anti-reduction lemma to get to this point, $W_6$ must also be constructed to be a private future world of $W_5$.

### 2018-05-13 toplas/paper.tex
Fix some latex bugs in the definition of Cauchy sequences and such.
reformulated a sentence

@@
   This amounts to showing that the closures generated by \texttt{g1} also respect the invariants of $W$.
-  The reasoning concerning variable $x$ follows that of \citet{Dreyer:jfp12}, so we will ignore it in the remainder of the proof and focus on the reasoning related to this low-level scenario.
-\lars{preceding sentence is unclear: what is $x$, what does ``this low-level scenario'' refer to ?}
+  We reason about the local variable $x$ in the closure in the same way \citet{Dreyer:jfp12} does it.
+  We ignore $x$ in the remainder this proof sketch to focus on the parts of the proof related to the setting of the capability machine.
 
@@
   \end{itemize}
-  We say that an o.f.e. $\left(X, \seqn{\nequal{n}}\right)$ is \emph{inhabited} if there
+  We say that an o.f.e. $\left(X, \seqn{\nequal[n]}\right)$ is \emph{inhabited} if there
   exists an element $x \in X$.
@@
   \begin{align*}
-    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal{k} x_n
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
   \end{align*}
@@
   \begin{align*}
-    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal{k} x_n.
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
   \end{align*}

### 2018-05-14 toplas/paper.tex
Change the discussion of efficient memory range clearing to discuss the "Efficient Tagged Memory" paper's insights and refer to it.

@@
 Ideally, such support would take the form of a highly-optimized instruction for erasing a large block of memory.
-However, from a discussion with the designers of the CHERI capability machine, we gather that it is not immediately clear whether and how such a primitive could be implemented efficiently in the CHERI context.
-\itoplassug{Maybe refer to some of the papers
-  reviewer A, esop mentions for stack clearing. See comment in tex file. }
+Recent work suggests that such a feature could perhaps be added to processors like CHERI, using a special hardware cache that tracks whether or not a memory location contains zero \citep{Joannou2017EfficientTaggedMemory}.
 % http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf

### 2018-05-14 toplas/paper.tex
Add a note to the awkward example proof sketch to point out where the  requirement for memory clearing turns up.

@@
   Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
+  Note also that arguing that the memory satisfies $W_2$ when we invoke the callback only works because we cleared the stack entirely (including the unused part) before the invocation.
+  Otherwise, it might contain local values that are only known to be safe in $W_1$, but for which we do not know that they will remain safe in the private future world $W_2$.
   This part corresponds to arguing about Hyp-Callee in the \texttt{scall} lemma.

### 2018-05-14 toplas/paper.tex
Disabled comments and colouring
Updated the new TOPLAS introduction

@@
 \definecolor{Blue}{RGB}{92,26,143}
-\newenvironment{toplas}
-    {\color{OliveGreen}}{}
+\newenvironment{toplas}{}{}
+%    {\color{OliveGreen}}{}
 \newenvironment{toplassug}
-    {\color{Blue}}{}
-
-\newcommand{\itoplas}[1]
-    {{\color{OliveGreen} #1}}
-\newcommand{\itoplassug}[1]
-    {{\color{Blue} #1}}
+    {\begin{comment}}{\end{comment}}
+
+\newcommand{\itoplas}[1]{#1}
+%    {{\color{OliveGreen} #1}}
+\newcommand{\itoplassug}[1]{}
+%    {{\color{Blue} #1}}
 
@@
 \begin{toplas}
-  This paper is an extension of a published conference
-  paper\citep[Skorstengaard:esop2018]. We have made minor improvements to
-  readability and completeness throughout the paper, but we highlight the
-  following changes:
+  This paper is an extension of a published conference paper\citep{Skorstengaard:esop18}.
+  We have made minor improvements to readability and completeness throughout the paper, but we highlight the following changes:
   \begin{itemize}
-  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provide informal intuition about how the logical relation machinery comes into play in this setting.
-  \item We have added a proof sketch for the \nameref{thm:ftlr} and one for Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
+  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provide informal intuition about how the logical relation machinery comes into play on a capability machine.
+  \item We have added a proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
   \item We have added figures that illustrate central parts of the calling convention.
   \item We have added a section on malloc.
-    In particular, we provide the specification for malloc used in the examples of the paper.
-  \item We have added a section on macro instructions.
-    All the macro instructions are described and the implementation of \texttt{scall} is provided.
+    Specifically, we provide the specification for the malloc used in the examples of the paper.
+  \item We have added a section on macro instructions with descriptions of all the macros and the implementation of \texttt{scall}.
   \item We have added a section on reasoning about programs that run on a capability machine.
+    This section explains how one reason about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
+    It also introduces a number of lemmas that prove recurring bits once and for all.
   \item We have added many details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
-    Many of the details we did not include in the paper are now in the appendix of the paper.
+    Further, many of the details we did still not include in the paper are now in the appendix of this paper.
   \end{itemize}

### 2018-05-14 toplas/paper.tex
Moved the syntax figure
Commented bit out

@@
 
-\begin{figure}[tb]
+\begin{figure}[hb]
 \[\arraycolsep=1.4pt
@@
 
-\begin{wrapfigure}{R}{.4\textwidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (7) {$\rwlx$};
-    \node[main node] (8) [below left of=7] {$\readwritel$};
-    \node[main node] (1) [below right of=7] {$\rwx$};
-    \node[main node] (2) [below right of=1] {$\exec$};
-    \node[main node] (3) [below right of=2] {$\entry$};
-    \node[main node] (4) [below left of=1] {$\readwrite$};
-    \node[main node] (5) [below right of=4] {$\readonly$};
-    \node[main node] (6) [below right of=5] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (7) edge (8)
-    (7) edge (1)
-    (8) edge (4)
-    (1) edge (2)
-    (2) edge (3)
-    (2) edge (5)
-    (3) edge (6)
-    (1) edge (4)
-    (4) edge (5)
-    (5) edge (6);
-
-    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
-    \node[main node] (2l) [below of=1l] {$\local$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (1l) edge (2l);
-  \end{tikzpicture}
-  \caption{Permission and locality hierarchy.}
-  \label{fig:perm-hier}
-\end{wrapfigure}
-
 We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
@@
 \itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}. We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
+
+\begin{wrapfigure}{R}{.4\textwidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (7) {$\rwlx$};
+    \node[main node] (8) [below left of=7] {$\readwritel$};
+    \node[main node] (1) [below right of=7] {$\rwx$};
+    \node[main node] (2) [below right of=1] {$\exec$};
+    \node[main node] (3) [below right of=2] {$\entry$};
+    \node[main node] (4) [below left of=1] {$\readwrite$};
+    \node[main node] (5) [below right of=4] {$\readonly$};
+    \node[main node] (6) [below right of=5] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (7) edge (8)
+    (7) edge (1)
+    (8) edge (4)
+    (1) edge (2)
+    (2) edge (3)
+    (2) edge (5)
+    (3) edge (6)
+    (1) edge (4)
+    (4) edge (5)
+    (5) edge (6);
+
+    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
+    \node[main node] (2l) [below of=1l] {$\local$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (1l) edge (2l);
+  \end{tikzpicture}
+  \caption{Permission and locality hierarchy.}
+  \label{fig:perm-hier}
+\end{wrapfigure}
 
@@
 For full details on the linking table, we refer to the technical
-appendix~\citep{technical_appendix}.} \itoplas{(Reviewer A, esop suggest explaining the
+appendix~\citep{technical_appendix}.} \itoplassug{(Reviewer A, esop suggest explaining the
   linking model)\\}

### 2018-05-14 toplas/paper.tex
Figure placement.
Typos and minor changes

@@
 
-\begin{figure}[htb]
+\begin{figure}[b]
   \begin{tikzpicture}[scale=.7, every node={scale=.7}]
@@
   matter what kind of region is used to mask as it is the region name we must
-  retain. This is why we allow $\revoked$ regions to be replaced by other
+  retain. This is why we allow revoked regions to be replaced by other
   regions in the two future world relations. This may seem unnecessary as the
@@
   want public future worlds to be able to prevent reinstation in private future
-  worlds. Private future worlds cannot revoke $\perma$ regions, so we cannot
-  allow public future worlds to us $\perma$ regions as a mask for a revoked
+  worlds. Private future worlds cannot revoke permanent regions, so we cannot
+  allow public future worlds to us permanent regions as a mask for a revoked
   region.
@@
   \begin{align*}
-  H^\nwl_A,  H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
   \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
   \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
+  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
+  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
+    H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
   H^\pwl_A \; s \; \hat{W} \defeq{} & \left\{\npair{\hs} \middle|
@@
         \right\} \union \{\npair[0]{\ms} \} \\
-  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
-  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
+    H^\nwl_A :{} & \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
   H^\nwl_A \; s \;\hat{W} \defeq{} & \left\{\npair{\hs} \middle|
@@
   In our definition of worlds, nothing prevents a world from having regions that are overlapping.
-  This is an issue for the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
-  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness as it prevents the regions from overlapping.
+  This might seem like an issue the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
+  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness condition on the world as it prevents the regions from overlapping.
   The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
@@
     Induction over $n$. By definition of $\stder(W)$, show
-    $\npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
-    \in \observations(W)$ assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
+    \[
+      \npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
+      \in \observations(W)
+    \]
+    assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
     $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
@@
 that a realistic implementation of malloc needs to rely on internal state that
-changes after every invocation, and relies on invariants on that state. We can
+changes after every invocation and relies on invariants on that state. We can
 only expect that malloc behaves according to its specification, if its internal
@@
 some of the macros we present next.
+
+\begin{figure}[tb]
+    \centering
+\begin{lstlisting}
+ $i_1$ = move r_t1 pc
+ $i_2$ = lea r_t1 $\var{ofc}$
+ $i_3$ = load r_stk r_t1
+ $i_4$ = pop pc
+\end{lstlisting}
+    \caption{The restoration code used in \texttt{scall}. The variable
+      $\var{ofc}$ is 5 which is the offset to the address where the old stack
+      pointer is stored on the stack.}
+    \label{fig:scall-rest-code}
+\end{figure}
 
@@
   \end{toplas}
-
-\end{figure}
-\begin{figure}[htb]
-  \begin{toplas}
-    \centering
-\begin{lstlisting}
- $i_1$ = move r_t1 pc
- $i_2$ = lea r_t1 $\var{ofc}$
- $i_3$ = load r_stk r_t1
- $i_4$ = pop pc
-\end{lstlisting}
-    \caption{The restoration code used in \texttt{scall}. The variable
-      $\var{ofc}$ is 5 which is the offset to the address where the old stack
-      pointer is stored on the stack.}
-    \label{fig:scall-rest-code}
-  \end{toplas}
 \end{figure}
@@
 % Anti-reduction lemma
-Our capability machine only allows the final memory of an execution to be
-observed, so naturally the correctness lemmas we prove in
-\sectionname~\ref{sec:examples} are statements about the memory in the halted
-configuration. In order to prove something about some part of the final memory,
-we establish an invariant for it as a region in a world and show that the
-initial configuration is in the $\observations$-relation w.r.t.\ that world. The
-$\observations$-relation says that if the initial configuration halts
-successfully, then the memory in the halting state must still respect a private
-future world of the initial world. It is, however, bothersome and error prone to
-try to argue about the entire execution in one go. Instead we want to reason
-modularly, in the sense that we only want to reason about parts of the execution
-at a time. To this end, we prove an anti-reduction lemma that essentially says
-that if we can show that an initial configuration, $\Phi$, steps to a
-configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then
-$\Phi$ is also in the observation relation.
+Our capability machine only allows the final memory of an execution to be observed, so naturally the correctness lemmas we prove in \sectionname~\ref{sec:examples} are statements about the memory in the halted configuration.
+In order to prove something about some part of the final memory, we establish an invariant for it as a region in a world and show that the initial configuration is in the $\observations$-relation w.r.t.\ that world.
+The invariants we are interested in must be permanent regions so that they are not revoked during execution.
+The $\observations$-relation says that if the initial configuration halts successfully, then the memory in the halting state must still respect a private future world of the initial world.
+It is, however, bothersome and error prone to try to argue about the entire execution in one go.
+Instead we want to reason modularly, in the sense that we only want to reason about parts of the execution at a time.
+To this end, we prove an anti-reduction lemma that essentially says that if we can show that an initial configuration, $\Phi$, steps to a configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then $\Phi$ is also in the observation relation.
 \begin{lemma}[Anti-reduction for $\observations$]
@@
 % FTLR
-The anti-reduction works well when we reason about the execution of known code,
+The anti-reduction works when we reason about the execution of known code,
 because we can execute it. When we want to reason about unknown code, the
@@
 
-In the common case \texttt{scall} is used to invoke untrusted code, and we will
+In the common case, \texttt{scall} is used to invoke untrusted code, and we will
 only have very general assumptions about that code, in which case we will use
@@
     \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
-    \item A adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
+    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
     \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
@@
   This amounts to showing that the closures generated by \texttt{g1} also respect the invariants of $W$.
-  We reason about the local variable $x$ in the closure in the same way \citet{Dreyer:jfp12} does it.
-  We ignore $x$ in the remainder this proof sketch to focus on the parts of the proof related to the setting of the capability machine.
-
-  The capability for the closure is a global enter capability that we call $c_{f4}$.
+  We reason about the local variable $x$ in the closure in the same way \citet{Dreyer:jfp12} does.
+  We ignore $x$ in the remainder of this proof sketch to focus on the parts of the proof related to the setting of the capability machine.
+
+  The capability for the generated closure is a global enter capability that we call $c_{f4}$.
   The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary, i.e.
-  $c_{f4} \in \stdvr(W')$ where $W'$ is $W$ with the regions added after executing \texttt{g1}.
+  $c_{f4} \in \stdvr(W')$ where $W'$ is $W$ with relevant regions added after executing \texttt{g1}.
   The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg$ and memory $\ms$ satisfies a world $W_1$ that is a private future world of $W'$, i.e.\ $\reg \in \stdrr(W_1)$ and $\memSat[]{\ms}{W_1}$.
@@
   For the Hyp-Cont case, we get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
-  At this point, we can use the anti-reduction lemma a final time to get to the point where we jump to the return pointer.
+  At this point, we can use the anti-reduction lemma one final time to get to the point where we jump to the return pointer.
 
@@
   This means that in $W_6$, we can reinstate every temporary or revoked region of $W_1$.
-  In the private future worlds we constructed, we only took transitions that, relative to $W_1$, are public.
-  The worlds we were given, $W_3$ and $W_5$ may have taken arbitrary public transitions, but this is no problem with respect to the public future world relation.
+  In the private future worlds we constructed, we only took transitions that are
+  public relative to $W_1$.
+  The worlds we were given, $W_3$ and $W_5$, may have taken arbitrary public transitions, but this is no problem with respect to the public future world relation.
   In $W_5$ there were regions to handle the stack.
@@
   constructed by us, while the bottom row of worlds are given. $W_6$ is constructed such
-  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
+  that $W_6 \futurestr W_5$ and $W_6 \futurewk W_1$.}
   \label{fig:worlds-in-pf-g1}
@@
 \subsection{Awkward example}
-\textbf{The region for $x$} The region $\iota_x$, is the region omitted from the proof sketch for the awkward example.
+\textbf{The region for variable $x$} The region $\iota_x$, is the region omitted from the proof sketch for the awkward example.
 \figurename~\ref{fig:x-reg} illustrates the transition system of $\iota_x$.

### 2018-05-15 toplas/paper.tex
Changes in frontmatter

@@
   \institution{Aarhus University}            %% \institution is required
-  \streetaddress{Street1 Address1}
+  \streetaddress{\AA{}bogade 34}
   \city{Aarhus}
-  \state{State1}
   \postcode{8210}
@@
   \institution{KU Leuven}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{-}
-  \state{-}
-  \postcode{-}
+  \streetaddress{Celestijnenlaan 200 A}
+  \city{Leuven}
+  \postcode{3001}
   \country{Belgium}
@@
 
-\maketitle
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002978.10002986.10002990</concept_id>
+<concept_desc>Security and privacy~Logic and verification</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10010940.10010992.10010998.10010999</concept_id>
+<concept_desc>Software and its engineering~Software verification</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+\ccsdesc[500]{Security and privacy~Logic and verification}
+\ccsdesc[500]{Software and its engineering~Software verification}
 
@@
 \end{abstract}
+
+\maketitle
 

### 2018-05-15 toplas/paper.tex
included acknowledgements

@@
 
-% \begin{acks}
-% This research was supported in part by the ModuRes Sapere Aude Advanced Grant
-% from The Danish Council for Independent Research for the Natural Sciences (FNU). 
-% Support for an STSM was received from COST Action EUTypes (CA15123).
-% Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
-% \end{acks}
+\begin{acks}
+This research was supported in part by the ModuRes Sapere Aude Advanced Grant
+from The Danish Council for Independent Research for the Natural Sciences (FNU). 
+Support for an STSM was received from COST Action EUTypes (CA15123).
+Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
+\end{acks}
 

### 2018-05-16 toplas/paper.tex
Made a tikz figure to replace the the world/capability picture

@@
 \usepackage{tikz}
-\usetikzlibrary{positioning}
+\usetikzlibrary{positioning,patterns,arrows,automata,decorations.markings}
 \usepackage{multirow,bigdelim}
@@
   \centering
-  \includegraphics{w11}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \scope
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+    \endscope
+    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+    \draw (-1.5,0) node {};
+    \draw (-1.5,15) node {};
+
+    \foreach \x in {0.5,1,...,4.5}
+    {
+      \shade[left color=blue!70,right color=white] (0,\x) rectangle (4.5,\x+.5);
+      \draw (0,\x) rectangle (4.5,\x+.5);
+    };
+    \foreach \x in {5,5.5,...,7}
+    {
+      \fill[color=red!55] (0,\x) rectangle (4.5,\x+.5);
+      \draw (0,\x) rectangle (4.5,\x+.5);
+    };
+    \foreach \x in {7.5,8,...,10}
+    {
+      \fill[color=green!85] (0,\x) rectangle (4.5,\x+.5);
+      \draw (0,\x) rectangle (4.5,\x+.5);
+    };
+
+    \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=0pt},yshift=0pt]
+    (4.6,1.5) -- (4.6,3) node (a) [black,midway] {};
+    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.3},yshift=0pt]
+    (-0.1,0.5) -- (-0.1,3) node (b) [black,midway,xshift=0.8cm] {};
+    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt},yshift=0pt]
+    (-0.1,3) -- (-0.1,4.5) node (c) [black,midway,xshift=0.8cm] {};
+    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt},yshift=0pt]
+    (-0.1,5) -- (-0.1,7) node (d) [black,midway,xshift=0.8cm] {};
+    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.6},yshift=0pt]
+    (-0.1,7.5) -- (-0.1,10.5) node (e) [black,midway,xshift=0.8cm] {};
+
+    \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
+    {\fill[black] circle(1.5pt);}}}}
+
+    \draw[capa] (-4,8.5) edge[-latex,postaction={decorate},in=180,out=30,densely dashed] (-0.4,9.3);
+
+    \draw[capa] (-4,6.8) edge[-latex,postaction={decorate},in=180,out=-30] (-0.4,6);
+    \draw[capa] (-4,4.55) edge[-latex,postaction={decorate},in=180,out=-30] (-0.4,3.75);
+
+    \draw[capa] (-4,2.05) edge[-latex,postaction={decorate},in=180,out=-30,densely dashed] (-0.4,1.25);
+
+    \draw[capa] (4.25,6.25) edge[-latex,postaction={decorate},in=0,out=0] (4.9,2.25);
+
+    \draw (-2,4.25) edge[color=red,thick] (-1,3.25)
+          (-1,4.25) edge[color=red,thick] (-2,3.25);
+    \draw (5.35,3.85) edge[color=red,thick] (6.35,2.85)
+          (6.35,3.85) edge[color=red,thick] (5.35,2.85);
+
+    \node at (-8,10) (perm) {\footnotesize Permanent region};
+    \node at (-8,9) (temp) {\footnotesize Temporary region};
+    \node at (-8,8) (temp) {\footnotesize Local capability};
+    \node at (-8,7) (temp) {\footnotesize Global capability};
+
+    \fill[color=green!85] (-12.5,9.75) rectangle (-11.5,10.25);
+    \fill[color=red!55] (-11.5,9.75) rectangle (-10.5,10.25);
+    \fill[color=green!85] (-11.75,9.75) -- (-11.75,10.25) -- (-11.25,10.25);    
+    \fill[color=red!55] (-11.75,9.75) -- (-11.25,9.75) -- (-11.25,10.25);
+    \draw (-12.5,9.75) rectangle (-10.5,10.25);
+
+    \shade[left color=blue!70,right color=white] (-12.5,8.75) rectangle (-11.5,9.25);
+    \draw (-12.5,8.75) rectangle (-10.5,9.25);
+
+    \draw[capa] (-12.4,8) edge[-latex,postaction={decorate},in=180,out=30,densely dashed] (-10.5,8);
+    \draw[capa] (-12.4,7) edge[-latex,postaction={decorate},in=180,out=30] (-10.5,7);
+    \draw (-13,10.6) rectangle (-5.4,6.4);
+  \end{tikzpicture}
 

### 2018-05-17 fully-abstract-comp/ftlr-proof-cases.tex
Added more cases to FTLR proof

@@
   \begin{itemize}
-  \item \item $\Phi_S \step \Phi_S'$
+  \item $\Phi_S \step \Phi_S'$
   \item $\Phi_T \step \Phi_T'$
@@
     \begin{itemize}
-    \item (restrict,cca,seta2b) $w_1 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $w_2 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\Phi_S(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\Phi_T(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\perm' \sqsubseteq\perm$,
-      $k = 1$
-    \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $w_2 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\Phi_S(r_1) = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$ and $\Phi_T(r_1) = ((\perm',\lin),\baddr,\eaddr,\aaddr')$ and $\perm' \sqsubseteq\perm$,
-      $k = 1$
-    \item ...
+    \item (restrict,cca,seta2b) $w_1 = ((\perm',\lin),\baddr,\eaddr,\aaddr')$, $w_1' = ((\perm',\lin),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi_T(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k = 1$
+    \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$, $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k = 1$
+    \item (move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S(r_1)$, and
+      $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg
+      \isLinear{w_1} \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$)
+      and $k = 2$.
+    \item (cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' = \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k=1$
+    \item (split) $\Phi_T(r_3) = \Phi_S(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$, $w_1 = w_1' = ((\perm,\lin),\baddr,n,\aaddr)$, $w_2 = w_2' = ((\perm,\lin),n+1,\eaddr,\aaddr)$, $k=2$.
+    \item (split) $\Phi_T(r_3) = \Phi_S(r_3) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \leq n$, $n < \sigma_\eaddr$, $w_1 = w_1' = \seal{\sigma_\baddr,n,\sigma}$, $w_2 = w_2' = \seal{n+1,\sigma_\eaddr,\sigma}$, $k=2$.
+    \item (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' = ((\perm,\linear),\baddr,n,\aaddr)$, $w_2 = \stkptr{\perm,n+1,\eaddr,\aaddr}$, $w_2' = ((\perm,\linear),n+1,\eaddr,\aaddr)$, $k=2$.
+    \item (splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
+      $\Phi_S(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_}$, and
+      $\Phi_T(r_3) = ((\perm,\linear),\eaddr_2+1,\eaddr_3,\aaddr_3)$,
+      $\Phi_S(r_3) = \stkptr{\perm,\eaddr_2+1,\eaddr_3,\aaddr_3}$, and
+      $\baddr_2 \leq \eaddr_2$, and
+      $\eaddr_2+1 \leq \eaddr_3$, and
+      $w_1=((\perm,\linear),\baddr_2,\eaddr_3,\aaddr_3)$, $w_1'=
+      \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_3}$, and
+      $w_2 = w_2' = w_3 = w_3' = 0$, and
+      $k=3$
+    \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
+      ((\perm,\lin),\baddr_2,\eaddr_2,\_)$, and
+      $\Phi_T(r_3) = \Phi_S(r_3) = ((\perm,\lin),\eaddr_2+1,\eaddr_3,\aaddr_3)$,
+      and
+      $\baddr_2 \leq \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
+      $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and
+      $\lin = \linear \Rightarrow w_2 = w_2' = w_3 = w_3' = 0$, and
+      $\lin = \normal \Rightarrow (w_2 = w_2' = \Phi_T(r_2) \text{ and } w_3 =
+      w_3' = \Phi_T(r_3))$, and $k=3$
+    \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) = \seal{\sigma_{\baddr,2},\sigma_{\eaddr,2},\_}$, and $\Phi_T(r_3) = \Phi_S(r_3) = \seal{\sigma_{\eaddr,2}+1,\sigma_{\eaddr,3},\sigma_{\aaddr,3}}$, and $\sigma_{\baddr,2} \leq \sigma_{\eaddr,2}$, and $\sigma_{\eaddr,2} + 1 \leq \sigma_{\eaddr,3}$and $w_1 = w_1' = \seal{\sigma_{\baddr,2}, \sigma_{\eaddr,3}, \sigma}$
       \end{itemize}
   \end{itemize}
-\item Other cases: store, load, jmp, jnz, xjmp
+\item All of the following hold: (includes memory-manipulation cases: store, load, that can be handled mostly uniformly)
+  \begin{itemize}
+  \item $\Phi_S \step \Phi_S'$
+  \item $\Phi_T \step \Phi_T'$
+  \item $\Phi_S' =
+    \updPcAddr{\Phi_S\updReg{r_1,r_2}{w_1,w_2}\update{\mem.\aaddr}{w}}$
+  \item $\Phi_T' =
+    \updPcAddr{\Phi_T\updReg{r_1',r_2'}{w_1',w_2'}\update{\mem.\aaddr}{w'}}$
+  \item One of the following hold:
+    \begin{itemize}
+    \item (store) $w_1 = w_1' = \Phi_S(r_1) = \Phi_T(r_1) =
+      ((\perm,\lin),\baddr,\eaddr,\aaddr)$, and $\perm \in \writeAllowed{}$,
+      and
+      $\withinBounds{w_1}$, and\\
+      $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $\isLinear{w} \Rightarrow
+      w_2 = w_2' = 0$, and $\neg \isLinear{w} \Rightarrow (w_2 = \Phi_S(r_2)
+      \text{ and } w_2' = \Phi_T(r_2))$.
+    \item (load) $w_2 = w_2' = \Phi_T(r_2) = \Phi_S(r_2) =
+      ((\perm,\lin),\baddr,\eaddr,\aaddr)$, and $\perm \in \readAllowed{}$,
+      $\withinBounds{((\perm,\lin),\baddr,\eaddr,\aaddr)}$, and \\
+      $w_1 = \Phi_S.\mem(\aaddr)$, and $w_1' = \Phi_T.\mem(\aaddr)$, and \\
+      $\isLinear{w_1} \Rightarrow w = w' = 0$, and \\
+      $\neg \isLinear{w_1} \Rightarrow w = w_1 \text{ and } w' = w_1'$
+    \end{itemize}
+  \end{itemize}
+\item All of the following hold: (includes memory-manipulation cases: store, load, that can be handled mostly uniformly (stack))
+  \begin{itemize}
+  \item $\Phi_S \step \Phi_S'$
+  \item $\Phi_T \step \Phi_T'$
+  \item $\Phi_S' =
+    \updPcAddr{\Phi_S\updReg{r_1,r_2}{w_1,w_2}\update{\ms_\stk.\aaddr}{w}}$
+  \item $\Phi_T' =
+    \updPcAddr{\Phi_T\updReg{r_1',r_2'}{w_1',w_2'}\update{\ms_\stk.\aaddr}{w'}}$
+  \item One of the following hold:
+    \begin{itemize}
+    \item (store) $w_1 = \Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $w_1' =  \Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \writeAllowed{}$,
+      and
+      $\withinBounds{w_1}$, and\\
+      $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $\isLinear{w} \Rightarrow
+      w_2 = w_2' = 0$, and $\neg \isLinear{w} \Rightarrow (w_2 = \Phi_S(r_2)
+      \text{ and } w_2' = \Phi_T(r_2))$.
+    \item (load) $ w_2' = \Phi_T(r_2) =
+      ((\perm,\linear),\baddr,\eaddr,\aaddr)$, and $w_2 = \Phi_S(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \readAllowed{}$,
+      $\withinBounds{((\perm,\lin),\baddr,\eaddr,\aaddr)}$, and $\aaddr \in
+      \dom(\Phi.\ms_\stk)$, and $\aaddr \in \dom(\Phi.\ms_\stk)$, and\\
+      $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and \\
+      $\isLinear{w_1} \Rightarrow w = w' = 0$, and \\
+      $\neg \isLinear{w_1} \Rightarrow w = w_1 \text{ and } w' = w_1'$
+    \end{itemize}
+  \end{itemize}
+
+\item All of the following hold: (includes memory-manipulation cases: jmp, jnz, xjmp, that can be handled mostly uniformly)
+  \begin{itemize}
+  \item 
+  \end{itemize}
 \end{itemize}

### 2018-05-22 fully-abstract-comp/ftlr-proof-cases.tex
Added the remaining cases for FTLR proof refactoring.

@@
 \[
-\npair{((\rx,\lin),\baddr,\eaddr,\aaddr),
-       ((\rx,\lin),\baddr,\eaddr,\aaddr)} \in \lreg{\trust}(W)
+\npair{(c,c)} \in \lreg{\trust}(W)
 \]
-We write $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
+For $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
 
@@
   \end{itemize}
-\item All of the following hold: (includes simple cases: gettype, geta, getb, gete, getp, getl, lt, plus, minus, that can be handled uniformly)
+\item All of the following hold: (includes simple cases: gettype, geta, getb,
+  gete, getp, getl, lt, plus, minus, one case of move that can be handled uniformly)
   \begin{itemize}
@@
 
-\item All of the following hold: (includes memory-manipulation cases: jmp, jnz, xjmp, that can be handled mostly uniformly)
+\item All of the following hold: (includes control-flow manipulation cases: jmp, jnz, xjmp, that can be handled mostly uniformly)
   \begin{itemize}
-  \item 
+  \item $\Phi_S \step \Phi_S'$
+  \item $\Phi_T \step \Phi_T'$
+  \item One of the following holds
+    \begin{itemize}
+    \item(jnz) $\Phi_S' = \updPcAddr{\Phi_S}$ and $\Phi_T' = \updPcAddr{\Phi_T}$
+    \item(jmp,jnz) $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and $\Phi_T' =
+      \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and 
+      $\isLinear{\Phi_S(r_1)} \Rightarrow w_1 = w_1' = 0$ and
+      $\neg \isLinear{\Phi_S(r_1)} \Rightarrow (w_1 = \Phi_S(r_1) \text{ and }
+      w_1' = \Phi_T(r_1))$
+    \item(xjmp) $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) =
+      \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and
+      $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\Phi_S'(r_1,r_2) =
+      \linCons{c_1},\linCons{c_2}$ and $\Phi_T'(r_1,r_2) =
+      \linCons{c_1'},\linCons{c_2'}$ and\\
+      $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
+        $\nonExec{c_\data}$ and $\Phi_T'(\pcreg,\rdata) = c_1',c_2'$ and
+      one of the following hold:
+      \begin{itemize}
+      \item $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$ and
+        $\nonExec{c_\data}$ and $\Phi_S'(\pcreg,\rdata) = c_1,c_2$.
+      \item $c_1 = \retptrc(\baddr,\eaddr,\aaddr)$, and $c_2 =
+        \retptrd(\aaddr_\stk,\eaddr_{\stk,\priv})$ and $\Phi(r_\stk) =
+        \stkptr{\rw,\stkb,\eaddr_\stk,\_}$ and $\Phi_S.\stk = \stk_\var{frame}
+        :: (\aaddr,\ms_{\stk,\priv})$ and
+        $\dom(\ms_{\stk,\priv} = [\eaddr_\stk+1,\eaddr_{\stk,\priv}])$ and
+        $\Phi_S'.\ms_\stk = \ms_{\stk,\priv} \uplus \Phi_S.\ms_\stk$ and
+        $\Phi_S'.\stk = \stk$ and
+        \[
+          \begin{split}
+\Phi_S' = \Phi_S&[\pcreg \mapsto ((\rx,\normal),\baddr,\eaddr,\aaddr)]\\
+            &[\rstk \mapsto
+            \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},e_\stk+1}] \\
+            &[\rdata,\rtmp{1},\rtmp{2},r_2 \mapsto 0, 0, 0, 0]
+          \end{split}
+        \]
+      \end{itemize}
+    \end{itemize}
   \end{itemize}

### 2018-05-22 fully-abstract-comp/ftlr-proof-cases.tex
Fixed some wrong assumptions in FTLR proof
Added a missing case
Add simple proofs for some minor lemmas.

@@
 
+\begin{lemma}[purePart is duplicable]
+  \label{lem:purePart-duplicable}
+ For all $W$, we have that $W = W \oplus \purePart{W}$.
+\end{lemma}
+\begin{proof}
+  Follows from the definition of $\purePart{}$ and $\oplus$.
+\end{proof}
+
 \begin{lemma}[purePart is idempotent]
-  For all $W$, $\mathit{purePart}( \purePart{W} ) = \purePart{W}$
+  For all $W$, $\mathit{purePart}( \purePart{W} ) = \purePart{W}$ and $\purePart{W} = \purePart{W} \oplus \purePart{W}$.
   \label{lem:purePart-idempotent}
@@
 \begin{proof}
-  Follows easily from the definition.
+  The first part follows easily from the definition.
+  The second statement follows from the first, together with Lemma~\ref{lem:purePart-duplicable}.
+\end{proof}
+
+\begin{lemma}[purePart is monotone]
+  \label{lem:purePart-mono}
+ For all $W'\future W$, we have that $\purePart{W'} \future \purePart{W}$.
+\end{lemma}
+\begin{proof}
+  Follows easily from the definition of $\purePart{}$ and $\future$.
+\end{proof}
+
+\begin{lemma}[Increasing authority is the future]
+  \label{lem:oplus-future}
+  For all $W$, we have that:
+  $W_1 \oplus W_2 \future W_1$
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
+
+\begin{lemma}[Purity is a thing of the past]
+  \label{lem:world-fut-purePart}
+  For all $W$, we have that $W \future \purePart{W}$.
+\end{lemma}
+\begin{proof}
+  Consequence of Lemmas~\ref{lem:purePart-duplicable}
+  and~\ref{lem:oplus-future}.
 \end{proof}
@@
 \begin{itemize}
-\item $\Phi_S' = \xjumpResult{c}{c}{\Phi_S}$
-\item $\Phi_T' = \xjumpResult{c}{c}{\Phi_T}$
+\item $\Phi_S' = \Phi_S\updReg{\pcreg}{c}$
+\item $\Phi_T' = \Phi_T\updReg{\pcreg}{c}$
 \end{itemize}
@@
     \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$, $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k = 1$
+    \item (cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$, $w_1' = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$, $\Phi_S(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$ and $\Phi_T(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$, $k = 1$
     \item (move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S(r_1)$, and

### 2018-05-22 fully-abstract-comp/technical-report.tex
Add simple proofs for some minor lemmas.

@@
   \label{lem:purePart-idempotemt}
-  For all $W$, we have that $\purePart{W} = \purePart{W} \oplus \purePart{W}$ and $\purePart{W} = \purePart{}(\purePart{W})$.
+  For all $W$, we have that $\purePart{W} = \purePart{}(\purePart{W})$ and $\purePart{W} = \purePart{W} \oplus \purePart{W}$.
 \end{lemma}
@@
 \end{lemma}
-\begin{proof}
-  Direct consequence of Lemmas~\ref{lem:purePart-duplicable}
-  and~\ref{lem:oplus-future}.
-\end{proof}
 

### 2018-05-22 fully-abstract-comp/ftlr-proof-cases.tex
Restructured ftlr proof document. Now part of technical report.
Added a standalone document for FTLR proof.
Fixed a typo.
Deleted obsolete part of FTLR proof.

--- 2018-05-22 fully-abstract-comp/ftlr-proof-cases.tex
@@
-\documentclass[a4paper]{article}
-\usepackage[a4paper]{geometry}
-\usepackage[utf8]{inputenc}
-
-\usepackage[svgnames]{xcolor}
-
-\input{preamble}
-
-\title{FTLR proof cases}
-
-\begin{document}
-
-\maketitle
-\section{Lemmas}
+\subsection{Lemmas}
 In this section, I have listed lemmas that seem to be necessary for the FTLR proof.
@@
 \begin{lemma}[$\lro$ closed under antireduction (generalised previous lemma)]
-  For all $\Phi_S$, $\Phi_S'$, $\Phi_T$, $\Phi_T'$, $j_S,J_T$, $n$, if
+  For all $\Phi_S$, $\Phi_S'$, $\Phi_T$, $\Phi_T'$, $j_S,j_T$, $n$, if
   \begin{itemize}
@@
 
-\section{FTLR proof}
+\subsection{FTLR proof}
 By induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
@@
 \end{itemize}
-
-Now assume that one of the configurations terminates etc.
-
-\subsubsection*{Assuming source machine terminates}
-First show the former of the two, assume $\Phi_S\term[i]$.
-Since $\Phi_S \neq \halted$, we have that $1 \leq i \leq n'$.
-
-We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c$, $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$, $\npair{(\baddr,\eaddr)}\in \readCond{\lin,W}$ and $\npair{(\baddr,\eaddr)}\in \xReadCond{\lin,W}$.
-From Lemma~\ref{lem:readcond-writecond-work}, using Lemma~\ref{lem:non-expansive}, $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ (from assumption~\ref{item:mem-ass}) and $W \oplus W_R \oplus W_M$ is well-defined (assumption~\ref{item:worlds-def}), we get (for any $a' \in [\baddr,\eaddr]$) that $\npair[n'-1]{(\ms_S(a'),\ms_T(a'))} \in \lrv(W')$ for some $W'$ such that $W_M = W' \oplus W_M'$.
-
-Since $\Phi_S \neq \halted$ and $\Phi_S\term[i]$, it must take a next execution step and $\Phi_S \not\step\failed$. 
-This is only possible if one of the following holds:
-\begin{itemize}
-\item $\callCond{\Phi_S,r_1,r_2,\aaddr}$ and
-  $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$ and 
-  $\exec{\Phi_S(\pcreg)}$.
-
-  In this case, we have for all $i = 0..\calllen-1$ that $\npair[n']{(\ms_S(\aaddr+i),\ms_T(\aaddr+i))} \in \lrv(W')$ for some $W'$, and also $\ms_S(\aaddr+i) \in \ints$, so by definition of $\lrv$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i)$.
-\item ($\src{(\neg\callCond{\Phi_S,r_1,r_2,\aaddr}}$ or $\src{\eaddr < \aaddr + \calllen-1) \tand}$) and
-  $\withinBounds{\Phi_S(\pcreg)}$ and $\exec{\Phi_S(\pcreg)}$.
-
-  In this case, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(W')$ for some $W'$, and also $\ms_S(\aaddr) \in \ints$ (since $\Phi_S \not\step\failed$), so by definition of $\lrv$, we have that $\ms_T(\aaddr) = \ms_S(\aaddr)$.
-\end{itemize}
-
-We now consider the first step of $\Phi_S \term[i]$ (we have already mentioned that $i \geq 1$).
-The possible instructions that could have been executed are:
-\begin{enumerate}
-\item $\tfail$
-\item $\thalt$
-\item $\tjmp{\trg{r_1}}$
-\item $\tjnz{\trg{r_1}}{\trg{\rn}}$
-\item $\tisptr{\trg{r_1}}{\trg{r_2}}$
-\item $\tgeta{\trg{r_1}}{\trg{r_2}}$
-\item $\tgetb{\trg{r_1}}{\trg{r_2}}$
-\item $\tgetp{\trg{r_1}}{\trg{r_2}}$
-\item $\tgetlin{\trg{r_1}}{\trg{r_2}}$
-\item $\tmove{\trg{r_1}}{\trg{\rn}}$
-\item $\tstore{\trg{r_1}}{\trg{r_2}}$
-\item $\tload{\trg{r_1}}{\trg{r_2}}$
-\item $\tcca{\trg{r_1}}{\trg{\rn}}$
-\item $\trestrict{\trg{r_1}}{\trg{r_2}}{\trg{\rn}}$
-\item $\tlt{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$
-\item $\tplus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$
-\item $\tminus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$
-\item $\tsetatob{\trg{r_1}}{\trg{r_2}}$
-\item $\txjmp{\trg{r_1}}{\trg{r_1}}$
-\item $\tcseal{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$
-\item $\tsplice{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$
-\item $\scall[o]{}{r_1}{r_2}$
-\end{enumerate}
-
-\paragraph{Case: $\protect\tfail$}
-Contradiction, because then $\Phi_S \step \failed \not\nstep[i-1] \halted$.
-
-\paragraph{Case: $\protect\thalt$}
-In this case $\Phi_S \step \halted$.
-We've seen above that $\ms_T(\aaddr) = \ms_S(\aaddr)$, so because $\decInstr{}$ is shared by source and target, we have that $\Phi_T \step\halted$.
-
-\paragraph{Case: $\protect\tjmp{\trg{r_1}}$}
-
-\paragraph{Case: $\protect\tjnz{\trg{r_1}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tisptr{\trg{r_1}}{\trg{r_2}}$}
-
-\paragraph{Case: $\protect\tgeta{\trg{r_1}}{\trg{r_2}}$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
-
-\paragraph{Case: $\protect\tgetb{\trg{r_1}}{\trg{r_2}}$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
-
-\paragraph{Case: $\protect\tgete{\trg{r_1}}{\trg{r_2}}\mid$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
-
-\paragraph{Case: $\protect\tgetlin{\trg{r_1}}{\trg{r_2}}$}
-Similar to the case for $\tisptr{\trg{r_1}}{\trg{r_2}}$.
-
-\paragraph{Case: $\protect\tmove{\trg{r_1}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tstore{\trg{r_1}}{\trg{r_2}}$}
-
-\paragraph{Case: $\protect\tload{\trg{r_1}}{\trg{r_2}}$}
-Since $\Phi_S\not\step\failed$, we have that $\Phi_S \step \Phi_S' \term[i-1]$ and one of the following cases hold: 
-\begin{itemize}
-\item \[\Phi_S' = \updPcAddr{\Phi\update{\mem.a}{w_2}\updReg{r_1}{w}}\]
-  with $\Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ 
-  $\perm \in \readAllowed{}$, $\withinBounds{\Phi(r_2)}$, $w = \Phi.\mem(a)$ and $w_2 = \linCons{w}$ and $\linConsPerm{w,\perm}$
-\item \[\Phi_S' = \updPcAddr{\Phi \update{\ms_\stk.a}{w_2}\updReg{r_1}{w}}\]
-  with $\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $\perm \in \readAllowed{}$, $\withinBounds{\Phi(r_2)}$, $\aaddr \in \dom(\Phi.\ms_\stk)$, $w = \Phi.\ms_\stk(a)$ and $w_2 = \linCons{w}$ and $\linConsPerm{w,\perm}$
-\end{itemize}
-
-\paragraph{Case: $\protect\tcca{\trg{r_1}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\trestrict{\trg{r_1}}{\trg{r_2}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tlt{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tplus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tminus{\trg{r_1}}{\trg{\rn}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tsetatob{\trg{r_1}}{\trg{r_2}}$}
-
-\paragraph{Case: $\scall[o]{}{r_1}{r_2}$}
-\dominique{Adapt to new treatment of $H_\sigma$ as non-global}
-In case $\scall[o]{}{r_1}{r_2}$ is the first step of the execution.
-We can assume $r_1,r_2 \neq \pcreg$ as this would cause the execution to fail\footnote{Assuming $\pcreg$ was one of the two registers, then the step relation would reqiure the register to contain an executable capability while the call semantics would require there to be a sealed capability.}\footnote{{\color{red} TODO:} We still have to consider what happens if the temp register is jumped to.}.
-By assumption \ref{item:reg-ass}, $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ which in particular means that for suitable $W_{R,1}$ and $W_{R,2}$
-\[
-  \npair[n']{(\src{\reg_S}(r_i),\reg_T(r_i))} \in \lrv(W_{r,1}) \text{ for i = 1,2}
-\]
-We can further deduce that $\src{\reg_S}(r_i)$ for $i=1,2$ are sealed capabilities as the execution would otherwise fail. This gives us
-\[
-  \begin{array}{l}
-    \forall W' \future W_{R,1}, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
-    \quad \npair[n'']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W' \oplus W_o) \wedge \dots
-  \end{array}
-\]
-and
-\[
-  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}.
-\]
-At this point, we need to pick a suitable $W' \future W_{R,1}$.
-To this end, we need to know a bit about $W_{R,1}$.
-From the operational semantics and in particular the semantics of $\scall{}{}{}{}$, we know $\Phi_S(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$ and by assumption $\npair[n']{(\src{\reg_S(\rstk)},\trg{\reg_T(\rstk)})} \in \lrv(W_{R,\stk})$.
-From this we can conclude that there is a stack region (specifically a spatially owned region) $r$ for which $\pwfree[W_{R,\stk}](r) \nequal[n'-1] \stdreg{[\baddr_\stk',\eaddr_\stk'],\gc}{\spa}$ where $[\baddr_\stk',\eaddr_\stk'] \supseteq [\baddr_\stk,\eaddr_\stk]$\footnote{Actually concluding this involves knowing that $W_{R,\stk}$ is ``well-formed'' which we use the heap satisfaction assumption to conclude.
-Perhaps a notion of well-formed worlds would be good along with a couple of lemmas that use it to make this reasoning once and for all.}.
-As $W_{R,\stk} \oplus W_{R,1}$ is well-formed, $\pwfree[W_{R,1}](r)$ must be a similar spatial region to the one in $W_{R,\stk}$.
-\dominique{How do we know $r_1 \neq \rstk$? Exclude in syntax?}
-We pick $W'$ to be
-\[
-  \begin{array}{r l}
-    W' = W     & \update{\mathrm{free}.r}{\revoked}\\
-               & \update{\mathrm{free}.x}{\stdreg{[\baddr_\stk',\aaddr_\stk -1],\gc}{\spa}}\\
-               & \update{\mathrm{priv}.y}{(\stdreg{[\aaddr_\stk,\eaddr_\stk],\gc}{\spa},\opc)}\\
-               & \update{\mathrm{free}.z}{\stdreg{[\eaddr_\stk+1,\eaddr_\stk'],\gc}{\spa}}
-  \end{array}
-\]
-\dominique{16-3-2018: shouldn't $W'$ be defined as a modification of $W_{R,1}$ rather than $W$?}
-\lau{12-03-2018: Write what $\opc$ is.}
-where $x$, $y$, and $z$ are fresh region names in the respective part of the world.
-To indicate similar worlds to $W'$ with the same ownership as another world well-defined w.r.t.
-$W$, we write $W'_{R,1}$ with the same spatially owned region as $W_{R,1}$.
-Using the private part of the future world relation, it is easy to see that $W'$ is a future world of $W_{R,1}$ (by the operational semantics $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$, so the new region at $y$ is non-empty).
-Using Lemma~\ref{lem:monotonicity}, we get
-\[
-  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W'_{R,2}
-\]
-which we use to get:
-\dominique{instantiate $n''$ to something?}
-\[
-  \npair[n'']{(\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2))} \in \lrexj(W_{R,1}' \oplus W_{R,2}')
-\]
-In order to use this, we need to pick a register-file pair and suitable memories.
-We pick a $\Phi_S$ in $\lrexj$, so the $\Phi_S'$ in $\lrexj$ becomes the $\Phi_S'$ we assumed $\Phi_S$ steps to.
-Specifically
-\[
-  \Phi_S.\reg
-  \begin{array}[t]{l}
-    \update{r_1}{\linCons{\Phi_S(r_1)}}\\
-    \update{r_2}{\linCons{\Phi_S(r_2)}}\\
-    \update{\pcreg}{0} \\
-    \update{\rdata}{0} \\
-    \update{\rstk}{c_\stk}\\
-    \update{\rretc}{c_{\ret,c}}\\
-    \update{\rretd}{c_{\ret,d}}\\
-    \update{\rtmp{1}}{0}\\
-  \end{array}
-\]
-where
-\begin{itemize}
-\item $c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}$
-\item $c_{\ret,c} = \retptrc(\sigma',(\perm,\lin),\baddr,\eaddr,\aaddr+\calllen))$
-\item $c_{\ret,d} = \retptrd(\sigma',(\aaddr_\stk,\eaddr_\stk))$
-\end{itemize}
-and
-$\ms_S,((\opc,\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}) :: \stk),\ms_\stk-\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}$
-
-For the target side, we pick the register-file
-\[
-  \Phi_T.\reg
-  \begin{array}[t]{l}
-    \update{r_1}{\linCons{\Phi_T(r_1)}}\\
-    \update{r_2}{\linCons{\Phi_T(r_2)}}\\
-    \update{\pcreg}{0} \\
-    \update{\rdata}{0} \\
-    \update{\rstk}{c_\stk'}\\
-    \update{\rretc}{c_{\ret,c}'}\\
-    \update{\rretd}{c_{\ret,d}'}\\
-    \update{\rtmp{1}}{0}\\
-  \end{array}  
-\]
-where
-\begin{itemize}
-\item $c_\stk' = ((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1)$
-\item $c_{\ret,c}' = \sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)}$
-\item $c_{\ret,d}' = \sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}$
-\end{itemize}
-\lau{We do not actually check the linearity of the stack, but it should have been checked that the base address actually is the stack base address which would give us that it is linear.
-This is not an assumption we have formally stated at the moment!}
-and the same memory
-\[
-  \Phi_T.\ms_T\update{\aaddr_\stk}{42}
-\]
-We have to argue (1) that this register-file pair is in the register-file relation for a world $W_R'$ with the same ownership as $W_R$ and regions $x$ and $y$ are owned, and (2) the memory-pair satisfies the memory satisfaction relation for a world $W_M'$ (i.e.
-with the same ownership as $W_M$).
-
-Let's start with (2).
-Here we use that we took a standard region and replaced by three standard regions.
-This means that when we split up the memory, we can reuse the same splitting of memory and world (w.r.t memory) except for the memory segment that previously satsfied the standard region in question.
-This memory segment, we split in three, so each piece fit with one of the new regions.
-
-For (1), we need to consider the contents of all registers.
-For the registers that were not overwritten, their safety follows from the register-file relation assumption we start with along with monotonicity and non-expansiveness of the value relation.
-This is also the case for $r_i$, $i=1,2$ if the capability is linear and it is trivial otherwise (as it is $0$).
-$\rtmp{1}$ is also trivial as it is $0$.
-This leaves $\rstk$, $\rretc$, and $\rretd$.
-
-For $\rstk$, we make sure the $x$ region is owned in the world we use.
-Now we use that we had a standard region governing the stack to get that everything on there is in the value relation.
-We then use the monotonicity of the value relation to conclude that everything is in the value relation in this future world which satisfies the read and write condition.
-
-For $\rretc$ we pick $W'$ as our world.
-It should be the case that $H_\sigma \; \sigma' \; W''_o$ only contains one \emph{executable} pair, namely the one we consider.
-Let $W'' \future W$, $n''<n'$, and $(c_S',\vsc_T') \in H_\sigma \; \sigma' \; W''_o$ be given and show
-\[
-  \npair[n'']{(c_{\ret,c},c_S',\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots)},\sealed{\sigma',\vsc_T})} \in \lrexj(W'')
-\]
-Here we can assume $(c_S',\vsc_T') \in H_\sigma \; \sigma' \;W''_o$ is a data capability pair\footnote{if it is not, then the execution would fail.
-Further, the other case we would have to consider ($c_{\reg,c}$ as the data capability) would also fail.}.\footnote{{\color{red}This suggest that $H_\sigma$ should accept only stack capabilities and specifically non-empty ones.}}
-
-{\color{DarkRed}
-  A this point, we would like to use the induction hypothesis with the assumptions we get from $\lrexj$. That is, we use the induction hypothesis to (hopefully) get
-
-\[
-  \npair[n'']{((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen-\dots),((\perm,\lin),\baddr,\eaddr,\aaddr+\calllen)} \in \lre(W'')
-\]
-(notice that the target has ``caught up'' here.) The assumptions we get from $\lrexj(W'')$ are not synced, so this would indicate that we need the target side to catch up.
-}
-
-\paragraph{Case: $\protect\txjmp{r_1}{r_2}$} 
-\dominique{Adapt to new treatment of $H_\sigma$ as non-global}
-In the case $\txjmp{r_1}{r_2}$\footnote{$\pcreg$ is a corner case here.
-It will fail contradicting previous assumption of termination.} was executed, we know that $\Phi_S(r_1)$ and $\Phi_S(r_2)$ either points to two sealed capabilities or a return pointer pair (all other cases result in $\Phi_S \step \failed$ contradicting $\Phi_S\term[i]$).
-From assumption \ref{item:reg-ass} we know
-\[
-  \npair[n']{(\Phi_S(r_i), \Phi_T(r_i))} \in \lrv(W_{R,i}) \text{ for $i = 1,2$}
-\]
-where $W_{R,i}$ are partitions of the partitioning of $W_R$ we get from assumption \ref{item:reg-ass}.
-From the previous observation w.r.t.
-$\Phi_S(r_1)$ and $\Phi_S(r_2)$, we know that we are in the "sealed capability case" of $\lrv$.
-This in particular gives us that
-\[
-  \begin{array}{l}
-    \forall W' \future W_{r,1}, W_o, n'' < n', (c_S,c_T) \in H_\sigma \; \sigma \; W_o \ldotp \\
-    \quad \npair[n'']{\Phi_S(r_1),c_S,\Phi_T(r_1),\sealed{\sigma,c_T}} \in \lrexj(W' \oplus W_o) \wedge \dots
-  \end{array}
-\]
-and
-\[
-  (\Phi_S(r_2), \vsc_{T,2}) \in H_\sigma \; \sigma \; W_{R,2}
-\]
-where $\Phi_T(r_2) = \sealed{\sigma,\vsc_{T,2}}$.
-Now use the former with $W'=W_{R,1}$, $W_o = W_{R,2}$, and $n'' = i-1$ to get
-\[
-  \npair[i-1]{\Phi_S(r_1),\Phi_S(r_2),\Phi_T(r_1),\Phi_T(r_2)} \in \lrexj(W_{R,1} \oplus W_{R,2}) 
-\]
-Using $\lrexj$, we want to ``start from'' configuration pair $(\Phi_S,\Phi_T)$ and end up in $(\Phi_S',\Phi_T')$ where
-\begin{itemize}
-\item $\Phi_S' = \xjumpResult{\Phi_S(r_1)}{\Phi_S(r_2)}{\Phi_S}$
-\item $\Phi_T' = \xjumpResult{\Phi_T(r_1)}{\Phi_T(r_2)}{\Phi_T}$
-\end{itemize}
-so we simply pick the same memory as in $\Phi_S,\Phi_T$ and the same register file but with $r_1$ and $r_2$ cleared if $\Phi(r_i)$ is linear, respectively.
-Using $W_M$ memory satisfaction follows from assumption \ref{item:mem-ass}.
-Using $W_R$ with all the ownership of regions owned by $W_{R,i}$, $i =1,2$ stripped, the register-file satisfaction follows from \ref{item:reg-ass}.
-Finally the $\npair[i-1]{[(\Phi_S(r_1),\Phi_T(r_1)),(\Phi_S(r_2),\Phi_T(r_2))]}\in\lrp(W_{R,1},W_{R,2})$ follows from the $\lrv$ assumption.
-Finally the consistency of the worlds we use follows easily.
-
-All in all, this allows us to conclude
-\[
-  \npair[i-1]{\Phi_S',\Phi_T'} \in \lro
-\]
-From the assumption $\Phi_S\term[i]$, we get $\Phi'_S\term[i-1]$ which we use with the above to conclude $\Phi_T' \term$.
-By the operational semantics of $\txjmp{r_1}{r_2}$, $\Phi_T \step \Phi_T'$ with which we conclude $\Phi_T \term$.
-
-% Now show
-% \[
-%   \Phi_T \term[i] \Rightarrow \Phi_S \term
-% \]
-
-\paragraph{Case: $\protect\tcseal{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$}
-
-\paragraph{Case: $\protect\tsplit{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}{\trg{\rn}}$}
-
-\paragraph{Case: $\protect\tsplice{\trg{r_1}}{\trg{r_2}}{\trg{r_3}}$}
-
-\subsubsection*{Assuming target machine terminates}
-\dominique{I suspect it would be more economical to not treat this case separately from the above (source machine terminates), but rather prove both results simultaneously by induction on the length of the terminating execution.}
-\paragraph{Case: }
-In the case  is the first step of the execution,
-
 \end{proof}
 
-\begin{lemma}[call lemma]
-
-\end{lemma}
-
-\section{TODO/Notes}
-\begin{itemize}
-\item Boundary crossing can occur on jump, xjump, call \emph{and} step!
-  \begin{itemize}
-  \item Make a lemma for updatePc that handles the latter once and for all. Maybe make a general lemma for all boundary crossings
-    \begin{itemize}
-    \item updatePc special case
-    \item argue about boundary crossing, trusted/untrusted, sanity conditions
-    \end{itemize}
-  \end{itemize}
-\item Lemma about call
-\end{itemize}
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "technical-report"
+%%% End:

### 2018-05-22 fully-abstract-comp/technical-report.tex
Restructured ftlr proof document. Now part of technical report.
Added a standalone document for FTLR proof.
Fixed a typo.
Deleted obsolete part of FTLR proof.
Remove some duplicate lemmas.

@@
 \usepackage{url}
+\usepackage{hyperref}
 
@@
 
+\begin{lemma}[purePart is duplicable]
+  \label{lem:purePart-duplicable}
+ For all $W$, we have that $W = W \oplus \purePart{W}$.
+\end{lemma}
+\begin{proof}
+  Follows from the definition of $\purePart{}$ and $\oplus$.
+\end{proof}
+
 \begin{lemma}[purePart is idempotent]
@@
 \end{lemma}
-
-\begin{lemma}[purePart is duplicable]
-  \label{lem:purePart-duplicable}
- For all $W$, we have that $W = W \oplus \purePart{W}$.
-\end{lemma}
+\begin{proof}
+  The first part follows easily from the definition.
+  The second statement follows from the first, together with Lemma~\ref{lem:purePart-duplicable}.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Follows easily from the definition of $\purePart{}$ and $\future$.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Consequence of Lemmas~\ref{lem:purePart-duplicable}
+  and~\ref{lem:oplus-future}.
+\end{proof}
 
@@
 
+\section{Proofs and lemmas}
+\input{ftlr-proof-cases}
+
 % \clearpage

### 2018-05-22 fully-abstract-comp/technical-report.tex
Make "World specifies component" predicate more compositional

@@
     \msdata[3] = \msdata[1] \uplus \msdata[2] \\
-    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}&
-    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}\\
+    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
+    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
     \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
@@
 \end{mathpar}
+
+\begin{lemma}[Linking preserves component well-formedness]
+  \label{lem:linking-well-formedness}
+  If $\vdash \var{comp}_1$ and $\vdash \var{comp}_2$, then $\vdash \var{comp}_1 \bowtie \var{comp}_2$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
 
@@
 \left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
-      \\
-      \begin{array}{l}
+      \begin{array}{l}
+        \dom(\code) \subseteq \ta \tand\\
         \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
@@
 \left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
-      \\
-      \begin{array}{l}
+      \begin{array}{l}
+        \dom(\code) \subseteq \ta \tand\\
         \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
@@
 
-\subsection{Fundamental Theorem of Logical Relations}
+\subsection{World specifies component}
 
@@
   \item $\mscode = \biguplus_i \code_i$
-  \item One of the following holds:
-    \begin{itemize}
-    \item $\dom(\mscode) \subseteq \ta$ and $\sigrets = \biguplus_i \sigrets_i$ and $\sigcloss = \biguplus_i \sigcloss_i$.
-    \item $\dom(\mscode) \mathrel{\#} \ta$ and $\sigrets = \emptyset$ and $\sigcloss = \biguplus_i \sigrets_i \uplus \biguplus_i \sigcloss_i$.
-    \end{itemize}
+  \item For all $i$, one of the following holds:
+    \begin{itemize}
+    \item $\dom(\code_i) \subseteq \ta$
+    \item $\dom(\code_i) \mathrel{\#} \ta$
+    \end{itemize}
+  \item $\sigrets = \biguplus_i \sigrets_i$
+  \item $\sigcloss = \biguplus_i \sigcloss_i$.
   \end{itemize}
@@
 
+\begin{lemma}[World specifies linked component if it specifies each separately]
+  \label{lem:world-specifies-linked}
+  If
+  \begin{itemize}
+  \item $W \vdash_\gc \var{comp}_0$
+  \item $W \vdash_\gc \var{comp}_0'$
+  \end{itemize}
+  Then $W \vdash \var{comp}_{0} \bowtie \var{comp}_{0}'$
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
+
+\subsection{Fundamental Theorem of Logical Relations}
 \begin{theorem}[FTLR]

### 2018-05-22 fully-abstract-comp/technical-report.tex
Fix broken reference.

@@
 \begin{lemma}[purePart is idempotent]
-  \label{lem:purePart-idempotemt}
+  \label{lem:purePart-idempotent}
   For all $W$, we have that $\purePart{W} = \purePart{}(\purePart{W})$ and $\purePart{W} = \purePart{W} \oplus \purePart{W}$.

### 2018-05-22 fully-abstract-comp/technical-report.tex
Prove a compatibility theorem for linking, fix an odd cyclicity in the definitions with an appropriate later in the definition.

@@
 
+\begin{lemma}[World specifies component pure]
+  \label{lem:world-specifies-comp-pure}
+  If $W \vdash_\gc \var{comp}_0$ then $\purePart{W} \vdash \var{comp}_{0}$.
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
+
 \subsection{Fundamental Theorem of Logical Relations}
@@
       W \vdash \var{comp} \tand\\
-      \text{For all } W' \future W \ldotp \text{If } \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'})\\
+      \text{For all } W' \future W \ldotp \text{If } \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
       \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
@@
   \begin{itemize}
-  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$
-  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$
+  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$
+  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$
   \end{itemize}
@@
   \begin{itemize}
-  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W_1\oplus W_2)$.
+  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W)$.
   \end{itemize}
 \end{lemma}
+\begin{proof}
+  First consider the case where one of the two pairs of components have a pair of ``main'' capabilities.
+  By the definition of $\lrcomp$, these capabilities have to be part of that components' export list.
+  It is then easy to see that the other component cannot have ``main'' capabilities (otherwise linking is not defined) and it is sufficient to prove the result for the underlying components (without the ``main'' capabilities).
+  Therefore, we can restrict ourselves to the case where both pairs of components are of the form $\var{comp}_0$, i.e. no ``main'' capabilities.
+
+  By Lemma~\ref{lem:world-specifies-linked}, we have that $W \vdash \var{comp}_1\bowtie \var{comp}_2$.
+
+  By definition of $\bowtie$, we have that the following hold:
+  \begin{align*}
+    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
+    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
+    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
+    \mscode[3] = \mscode[1] \uplus \mscode[2] &
+    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
+    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
+    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
+    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
+    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
+  \end{align*}
+
+  Now take $W' \future W$ and assume that $\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
+  Then it remains to show that $\overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$.
+  We do this by complete induction on $n$, so that we can assume that $\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
+
+  For each $$\npair{(c_{\mathrm{export}},c_{\mathrm{export}})} \in \overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}}$$, we have that $\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}$ is either in $\overline{\npair{(c_{\mathrm{export},1},c_{\mathrm{export},1})}}$ or $\overline{\npair{(c_{\mathrm{export},2},c_{\mathrm{export},2})}}$.
+  By unfolding $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$, it suffices to prove that $$\overline{\npair[n']{(c_{\mathrm{import},1},c_{\mathrm{import},1})}} \in \lrv(\purePart{W'})$$ and $$\overline{\npair[n']{(c_{\mathrm{import},2},c_{\mathrm{import},2})}} \in \lrv(\purePart{W'})$$
+  for all $n' < n$.
+
+  Simple set algebra tells us that 
+  $$\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}} \subseteq \overline{c_{\mathrm{export},3}} \cup \overline{c_{\mathrm{import},3}}$$
+  and for all $n' < n$, we know that $$\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$$
+  and
+  $$\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})\text{.}$$
+\end{proof}
 

### 2018-05-23 fully-abstract-comp/technical-report.tex
Fix usage of the term program.

@@
 
-A program needs access to the following:
+A component needs access to the following:
 \begin{description}
@@
 \lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}
+
+\subsection{Reasonable components}
+\label{sec:reas-comp}
 \begin{definition}
@@
 \begin{definition}
-  A component $\program = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
-  \begin{itemize}
-  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $A$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-  \item $\plug{\context}{\program}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\plug{\context}{\program}$ handle trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+  A component $\var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
+  \begin{itemize}
+  \item $\plug{\context}{\var{comp}}$ checks $\stkb$ before calls in $A$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\plug{\context}{\var{comp}}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\plug{\context}{\var{comp}}$ handle trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\plug{\context}{\var{comp}}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
   \end{itemize}
@@
 
+\subsection{Full abstraction: precise formulation}
+\label{sec:full-abstr-prec}
+
 \begin{definition}[Source language contextual equivalence]
-  In the source language, we define that $\src{\program_1 \sconeq \program_2}$ iff
+  In the source language, we define that $\src{\var{comp}_1 \sconeq \var{comp}_2}$ iff
   \begin{equation*}
-    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\program_2} \sterm{\ta[,2],\stkb_2}}
+    \forall \src{\context} \ldotp \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
   \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\program_i.\mscode)}$
+  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$
 \end{definition}
@@
 \begin{definition}[Target language contextual equivalence]
-  In the target language, we define that $\trg{\program_1 \tconeq \program_2}$ iff
+  In the target language, we define that $\trg{\var{comp}_1 \tconeq \var{comp}_2}$ iff
   \begin{equation*}
-    \forall \trg{\context} \ldotp \trg{\plug{\context}{\program_1} \term} \Leftrightarrow \trg{\plug{\context}{\program_2} \term}
+    \forall \trg{\context} \ldotp \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
   \end{equation*}
@@
   \label{thm:full-abstraction}
-  For reasonable components $\program_1$ and $\program_2$, we have
+  For reasonable components $\var{comp}_1$ and $\var{comp}_2$, we have
   \begin{gather*}
-    \src{\program_1} \sconeq \src{\program_2}\\
+    \src{\var{comp}_1} \sconeq \src{\var{comp}_2}\\
     \Updownarrow\\
-    \comp{\src{\program_1}} \tconeq \comp{\src{\program_2}}
+    \comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}
   \end{gather*}

### 2018-05-23 fully-abstract-comp/technical-report.tex
Make a start with the "FTLR for components", fix a variety of minor problems here and there.

@@
     \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
+    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}}
   }{
@@
       \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
+      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; W \text{ for all $n' < n$}\wedge\\
       \quad (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
       \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad\forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
       \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
@@
         \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
-        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
-        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W} \wedge\\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
         % we are excluding rwx pointers.
@@
         \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
-                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
+                                  \npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}\endarray
       \end{array}
@@
       \perm \sqsubseteq \rx \wedge \\
-      \npair{(\baddr,\eaddr)} \in \xReadCond[\square,(\ta,\stkb)]{\normal,W}
+      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,(\ta,\stkb)]{\normal,W}
     \end{array}
@@
 \[
-  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
+  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
     \begin{array}{l}
@@
       \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
       \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
@@
 \[
-  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
+  \xReadCond{\lin,W} = \left\{ \npair{A} \middle| 
     \begin{array}{l}
@@
       \quad \exists M : S \fun \powerset{\MemSeg} \\
-      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
+      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
@@
 \[
-  \writeCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
+  \writeCond{\lin,W} = \left\{ \npair{A} \middle| 
     \begin{array}{l}
@@
       \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
       \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
@@
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-    \dom(\code) \subseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
+    \dom(\code) \subseteq \ta \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset) \wedge \\
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
@@
   \begin{itemize}
-  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\pure}$
+  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\lin}$ with $\lin = \pure$ or $\lin = \spatialo$
   \item $\dom(\msdata) = \biguplus_i A_i$
@@
   \begin{itemize}
-  \item $\trust = \trusted \vee \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
-  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
+  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}$
   \end{itemize}
@@
      \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
-      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
-     \{(c_{\mathrm{main},c}, c_{\mathrm{main},c}),(c_{\mathrm{main},d}, c_{\mathrm{main},d})\} \subseteq (\overline{c_{\mathrm{export}}},\overline{c_{\mathrm{export}}})
+     \{c_{\mathrm{main},c},c_{\mathrm{main},d}\} \subseteq \overline{c_{\mathrm{export}}}
        \end{multlined}
@@
 
+\begin{lemma}[FTLR for component data-values]
+  If
+  \begin{equation*}
+    \dom(\mscode), A_{\mathrm{own}}, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+  \end{equation*}
+  and
+  \begin{itemize}
+  \item $\npair{A_{\mathrm{non-linear}}} \in \readCond{\normal,W}$
+  \item $\npair{A_{\mathrm{non-linear}}} \in \writeCond{\normal,W}$
+  \item $\npair{A_{\mathrm{own}}} \in \readCond{\linear,W}$
+  \item $\npair{A_{\mathrm{own}}} \in \writeCond{\linear,W}$
+  \item $W(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
+  \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(W)$ for all $n' < n$
+  \item $w \not\in \overline{c_{\mathrm{import}}}$
+  \end{itemize}
+
+  then $\npair{(w,w)} \in \lrv(W)$
+\end{lemma}
+\begin{proof}
+  By induction on the judgement 
+  \begin{equation*}
+    \dom(\mscode), A_{\mathrm{own}}, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+  \end{equation*}
+  We have the following cases:
+  \begin{itemize}
+  \item $w = z$: result follows trivially
+  \item $w = c$, $c \in \overline{c_{\mathrm{import}}}$: contradicts assumption.
+  \item $w = ((\perm,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq A_{\mathrm{code}}$, $\perm \sqsubseteq \rx$:
+    TODO
+  \item $w = ((\perm,\lin),\baddr,\eaddr,\aaddr)$,
+    $\permbnf \sqsubseteq \rw$,
+    $\lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}}$ and
+    $\lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}$:
+    TODO
+
+  \item $w = \sealed{\sigma,\vsc}$ and
+    \begin{equation*}
+      A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc
+    \end{equation*}
+    and $\sigma \in \sigcloss$:
+
+    By definition of $\lrv$ and by choosing island $r = r_{\mathrm{code}}$, it suffices to prove that
+    \begin{itemize}
+    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; W$ for all $n' < n$:
+
+     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that
+     \begin{align*}
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc)} \in \lrv \; \xi(\hat{W})) \tor\\
+        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\hat{W}))
+     \end{align*}
+
+     TODO
+
+     this follows
+     by induction and Lemma~\ref{lem:non-expansive} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption, we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+   \item $(\isLinear{\vsc} \text{ iff } \isLinear{\vsc})$: trivially fine
+
+   \item If $\isLinear{\vsc}$, then 
+     for all $W' \future W$, $W_o$, $n' < n$ and $\npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that 
+     \begin{equation*}
+       \npair[n']{(\vsc,\vsc_S',\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o)) \wedge \\
+     \end{equation*}
+     TODO
+   \item 
+       If $\nonLinear{\src{\vsc_S}}$ then for all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+       \begin{equation*}
+         \npair[n']{(\vsc,\src{\vsc_S'},\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
+       \end{equation*}
+       TODO
+    \end{itemize}
+  \end{itemize}
+\end{proof}
+
 \begin{lemma}[FTLR for components]
@@
   \item $\gc = (\ta,\stkb)$
+  \item $\var{comp}$ is a well-formed component, i.e. $\vdash \var{comp}$
   \item One of the following holds:
     \begin{itemize}
-    \item $\trust = \trusted$, $\var{comp}$ is a reasonable component (see
-      Section~\ref{sec:full-abstraction}) and $\dom(\comp.\mscode) \subseteq \ta$
-    \item $\trust = \untrusted$ and $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
+    \item $\var{comp}$ is a reasonable component (see
+      Section~\ref{sec:reas-comp}) and $\dom(\var{comp}.\mscode) \subseteq \ta$
+    \item $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
     \end{itemize}
@@
 \end{lemma}
+\begin{proof}
+  If $\var{comp}$ is of the form $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$, then we have (by definition of component well-formedness) that $c_{\mathrm{main},c}, c_{\mathrm{main},d} \subseteq \var{comp}_0.\overline{c_{\mathrm{export}}}$, as required by $\lrcomp(W)$.
+  Hence, we can restrict our attention to components of the form $\var{comp}_0$.
+  Take
+  \begin{equation*}
+    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+  \end{equation*}
+
+  We know that $W \vdash \var{comp}$ by assumption, so it remains to prove that 
+  for all  $W' \future W$ with 
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
+  \end{equation*}
+  we have that
+  \begin{equation*}
+    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+  \end{equation*}
+
+  TODO
+\end{proof}
 Note: the trusted case of the above lemma can be considered as a compiler correctness result.
@@
 
-%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
+%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}
 \begin{definition}[Never store link seals and code capability]
@@
 \begin{definition}
-  A component $\var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
+  A component $\var{comp})$ with $\dom(\var{comp}.\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
   \begin{itemize}

### 2018-05-23 fully-abstract-comp/ftlr-proof-cases.tex
Remove some duplicate lemmas.
fixed mistake in ftlr proof (related to wrong E-rel def used)
Minor changes.

--- 2018-05-22 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \begin{itemize}
-  \item If $\npair{(b,e)} \in \readCond{\lin,W}$, then $\npair[n']{(b,e)} \in \readCond{\lin,W}$.
-  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
+  \item If $\npair{A} \in \readCond{\lin,W}$, then $\npair[n']{A} \in \readCond{\lin,W}$.
+  \item If $\npair{A} \in \stackReadCond{\lin,W}$, then $\npair[n']{A} \in \stackReadCond{\lin,W}$.
+  \item If $\npair{A} \in \writeCond{\lin,W}$, then $\npair[n']{A} \in \writeCond{\lin,W}$.
+  \item If $\npair{A} \in \stackWriteCond{\lin,W}$, then $\npair[n']{A} \in \stackWriteCond{\lin,W}$.
+  \item If $\npair{A} \in \execCond{\lin,W}$, then $\npair[n']{A} \in \execCond{\lin,W}$.
+  \item If $\npair{A} \in \xReadCond{\lin,W}$, then $\npair[n']{A} \in \xReadCond{\lin,W}$.
+  \item If $\npair{(w,w')} \in \lrvg{\trust}{W}$, then $\npair[n']{(w,w')} \in \lrvg{\trust}{W}$.
+  \item If $\npair{(w,w')} \in \lrrg{\trust}(W)$, then $\npair[n']{(w,w')} \in \lrrg{\trust}(W)$.
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}(W)$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}(W)$.
   \end{itemize}
@@
 
-\begin{lemma}[$\lrol$ closed under target language antireduction]
+\begin{lemma}[$\lrol$ closed under source language antireduction]
   For all $\Phi_S$, $\Phi_T$, $\Phi_T'$, $j$, $n$, if
@@
   From this, it follows that $\memSatFStack{\ms_\stk\update{a}{0},\ms_\var{T,heap}\update{a}{0}}{W_r' \oplus W_\var{free\_stack}'}$ and finally $\memSat{(\ms_S,\stk,\ms_\stk\update{a}{0},\ms_T\update{a}{0})}{W_M'}$.
-\end{proof}
-
-\begin{lemma}[purePart is idempotent]
-  For all $W$, $\mathit{purePart}( \purePart{W} ) = \purePart{W}$
-  \label{lem:purePart-idempotent}
-\end{lemma}
-\begin{proof}
-  Follows easily from the definition.
 \end{proof}

### 2018-05-23 fully-abstract-comp/technical-report.tex
Minor changes.

@@
       \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; W \text{ for all $n' < n$}\wedge\\
+      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
       \quad (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
@@
         \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
-        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackReadCond{W} \wedge \\
-        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \stackWriteCond{W}
+        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
       \end{array}
@@
 \[
-  \stackReadCond{W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
+  \stackReadCond{W} = \left\{ \npair{A} \middle| 
     \begin{array}{l}
@@
       \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
       \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
@@
 \[
-  \stackWriteCond{W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
+  \stackWriteCond{W} = \left\{ \npair{[\baddr,\eaddr]}) \middle| 
     \begin{array}{l}
@@
 \[
-  \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
+  \execCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)} \middle|
     \begin{array}{l}
@@
       \quad ((W_l = \purePart{W} \wedge \lin = \normal) \vee (W_l = W \wedge \lin = \linear)) \wedge\\
-      \quad \forall n' < n, W' \future W_l, \perm \in S_\perm, \aaddr \in [\baddr',\eaddr'] \subseteq [\baddr,\eaddr] \ldotp\\
-      \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr',\eaddr',\aaddr)}{((\perm,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+      \quad \forall n' < n, W' \future W_l, \aaddr \in [\baddr',\eaddr'] \subseteq [\baddr,\eaddr] \ldotp\\
+      \qquad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
     \end{array}
@@
   \item $W(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
-  \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
+  \item $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$
   \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(W)$ for all $n' < n$
@@
   \item $w = ((\perm,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq A_{\mathrm{code}}$, $\perm \sqsubseteq \rx$:
+    By definition of $\lrv(W)$, it suffices to prove that $\npair{[\baddr,\eaddr]}\in\readCond{\normal,W}$, $\npair{[\baddr,\eaddr]}\in\execCond{\normal,W}$ and $\npair{[\baddr,\eaddr]}\in\xReadCond{\normal,W}$.
+
+    \dominique{TODO: readCond probably not fine: the code of a component may contain stuff that must remain private.
+      We have to prevent that a component exports capabilities with authority over its code memory (directly or indirectly)?}
     TODO
@@
     \begin{itemize}
-    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; W$ for all $n' < n$:
+    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
 
      By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that
-     \begin{align*}
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc)} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\hat{W}))
-     \end{align*}
-
-     TODO
-
-     this follows
-     by induction and Lemma~\ref{lem:non-expansive} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption, we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+     one of the following holds: 
+     \begin{itemize}
+     \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrv \; \xi(\xi^{-1}(W))$
+     \item $(\dom(\code) \subseteq \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\xi^{-1}(W))$
+     \end{itemize}
+     Since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$ in both cases.
+
+     This follows by induction and Lemma~\ref{lem:non-expansive} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+
    \item $(\isLinear{\vsc} \text{ iff } \isLinear{\vsc})$: trivially fine
@@
      \end{equation*}
-     TODO
+     \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
    \item 
@@
        \end{equation*}
-       TODO
+       \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
     \end{itemize}

### 2018-05-23 fully-abstract-comp/technical-report.tex
Add a note about small problem in FTLR for components proof
Solve the issue with components exporting private code stuff, bit of work on proof.

@@
   \inference{
-    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} & \perm \sqsubseteq \rx
+    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} & 
+    \sigma \in \sigcloss
   }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\normal),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,((\perm,\rx),\baddr,\eaddr,\aaddr)}
   }
@@
 
+\begin{lemma}[FTLR for component code-values]
+  If $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} w$
+  and
+  \begin{itemize}
+  \item $W(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
+  \end{itemize}
+
+  then $\npair{(w,w)} \in \lrvg{\trust}(W)$.
+\end{lemma}
+\begin{proof}
+  By induction on the judgement $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} w$.
+
+  TODO
+\end{proof}
+
 \begin{lemma}[FTLR for component data-values]
@@
   \item $w = c$, $c \in \overline{c_{\mathrm{import}}}$: contradicts assumption.
-  \item $w = ((\perm,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq A_{\mathrm{code}}$, $\perm \sqsubseteq \rx$:
-    By definition of $\lrv(W)$, it suffices to prove that $\npair{[\baddr,\eaddr]}\in\readCond{\normal,W}$, $\npair{[\baddr,\eaddr]}\in\execCond{\normal,W}$ and $\npair{[\baddr,\eaddr]}\in\xReadCond{\normal,W}$.
-
-    \dominique{TODO: readCond probably not fine: the code of a component may contain stuff that must remain private.
-      We have to prevent that a component exports capabilities with authority over its code memory (directly or indirectly)?}
+  \item $w = \sealed{\sigma,((\perm,\rx),\baddr,\eaddr,\aaddr)}$, $[\baddr,\eaddr]\subseteq A_{\mathrm{code}}$, $\sigma \in \sigcloss$:
+
     TODO
@@
   TODO
+  \dominique{Something missing: we should be able to subdivide the world according to how linear authority is distributed in the component well-formedness judgement, but the world-specifies-component relation does not give us this.}
 \end{proof}

### 2018-05-24 fully-abstract-comp/ftlr-proof-cases.tex
Corrected name of downwards closure lemma

--- 2018-05-22 fully-abstract-comp/ftlr-proof-cases.tex
@@
 In this section, I have listed lemmas that seem to be necessary for the FTLR proof.
-\begin{lemma}[non-expansiveness of relations]
-  \label{lem:non-expansive}
+\begin{lemma}[Downwards closure of relations]
+  \label{lem:downwards-closed}
   If $n' \leq n$, then
   \begin{itemize}
-  \item If $\npair{(b,e)} \in \readCond{\lin,W}$, then $\npair[n']{(b,e)} \in \readCond{\lin,W}$.
+  \item If $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$, then $\npair[n']{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$.
+  \item If $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$, then $\npair[n']{(\baddr,\eaddr)} \in \readCond{\lin,W}$.
   \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
+  \item If $\npair{(\reg_S,\reg_T)} \in \lrr(W)$, then $\npair[n']{(\reg_S,\reg_T)} \in \lrr(W)$.
+  \end{itemize}
+\end{lemma}
+
+\begin{lemma}[Non-expansiveness of relations]
+  \label{lem:non-expansiveness}
+  If $W \nequal W'$, then
+  \begin{itemize}
+  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \nequal \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W'}$.
+  \item ...
   \end{itemize}
@@
     From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
-    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:non-expansive}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:downwards-closed}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
 
-  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
-    The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by non-expansiveness of memory satisfaction, i.e.
-Lemma~\ref{lem:non-expansive}.
+  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?}
+    and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
+    The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by downwards closure of memory satisfaction, i.e.~Lemma~\ref{lem:downwards-closed}.
   \end{itemize}
@@
     From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm = \perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrvg{\trust}(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \stackWriteCond{W}$.
-    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:non-expansive}, we then get that $\memSat[n']{(\ms_S,\stk,\ms_\stk\update{\aaddr}{0},\ms_T\update{\aaddr'}{0})}{W_M'}$.
+    From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:downwards-closed}, we then get that $\memSat[n']{(\ms_S,\stk,\ms_\stk\update{\aaddr}{0},\ms_T\update{\aaddr'}{0})}{W_M'}$.
 
-  \item case $\neg\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = \ms_\stk(\aaddr)$ and it follows from $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
-    The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by non-expansiveness of memory satisfaction, i.e.
-Lemma~\ref{lem:non-expansive}.
+  \item case $\neg\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = \ms_\stk(\aaddr)$ and it follows from $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?}
+    and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
+    The fact that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$ then follows simply by downwards closure of memory satisfaction, i.e.\ Lemma~\ref{lem:downwards-closed}.
   \end{itemize}
@@
 \end{itemize}
-and pick
-\begin{itemize}
-\item $\Phi_S' = \Phi_S\updReg{\pcreg}{c}$
-\item $\Phi_T' = \Phi_T\updReg{\pcreg}{c}$
-\end{itemize}
 and show
@@
     \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$, $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k = 1$
-    \item (cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$, $w_1' = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$, $\Phi_S(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$ and $\Phi_T(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma,\sigma'}$, $k = 1$
+    \item (cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $w_1' = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $\Phi_S(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$ and $\Phi_T(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $k = 1$
     \item (move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S(r_1)$, and

### 2018-05-24 fully-abstract-comp/ftlr-proof-cases.tex
Updated FTLR proof refactoring so it is easier to refer to specific cases.

@@
 \begin{lemma}[$\lro$ closed under antireduction (generalised previous lemma)]
+  \label{lem:lro-anti-red-gen}
   For all $\Phi_S$, $\Phi_S'$, $\Phi_T$, $\Phi_T'$, $j_S,j_T$, $n$, if
@@
 \begin{itemize}
-\item $\Phi_S = (\ms_S,\reg_S\updReg{\pcreg}{c},\stk,\ms_\stk)$
-\item $\Phi_T = (\ms_T,\reg_T\updReg{\pcreg}{c})$
+\item $\Phi_S = (\ms_S,\reg_S\update{\pcreg}{c},\stk,\ms_\stk)$
+\item $\Phi_T = (\ms_T,\reg_T\update{\pcreg}{c})$
 \end{itemize}
@@
 \[
-  \npair[n']{(\Phi_S',\Phi_T')} \in \lro
+  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
 \]
 In order to do this, we first, we prove that one of the following holds:
-\begin{itemize}
-\item $\Phi_S \step \failed$ and $\Phi_T \step\failed$
-\item $\Phi_S \step \halted$ and $\Phi_T \step\halted$
+\lau{TODO: add $\neq$ failed to the cases that are not failed.}
+\begin{enumproof}
+  \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T \step\failed$
+  \item $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
   \item All of the following hold:
@@
     \item $\exec{\Phi_S(\pcreg)}$
-  \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in \ints$.
+    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in
+      \ints$.
     \item TODO: other useful info
     \end{itemize}
-\item All of the following hold: (includes simple cases: gettype, geta, getb,
-  gete, getp, getl, lt, plus, minus, one case of move that can be handled uniformly)
+  \item \label{case:ftlr:get-and-arith-op} All of the following hold: (includes simple cases: gettype, geta, getb,
+    gete, getp, getl, lt, plus, minus, one case of move that can be handled
+    uniformly)
     \begin{itemize}
-  \item $\Phi_S \step \Phi_S'$
+    \item $\Phi_S \step[\gc] \Phi_S'$
     \item $\Phi_T \step \Phi_T'$
@@
     \end{itemize}
-\item All of the following hold: (includes cap-manipulation cases: move, cca, restrict, seta2b, cseal, split, splice, that can be handled mostly uniformly)
+  \item All of the following hold: (includes cap-manipulation cases: move, cca,
+    restrict, seta2b, cseal, split, splice, that can be handled mostly
+    uniformly)
     \begin{itemize}
-  \item $\Phi_S \step \Phi_S'$
+    \item $\Phi_S \step[\gc] \Phi_S'$
     \item $\Phi_T \step \Phi_T'$
-  \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r_1\cdots r_k}{w_1\cdots w_k}}\neq \failed$
-  \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r_1\cdots r_k}{w_1'\cdots w_k'}} \neq \failed$
+    \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r_1\cdots r_k}{w_1\cdots w_k}}\neq
+      \failed$
+    \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r_1\cdots r_k}{w_1'\cdots w_k'}}
+      \neq \failed$
     \item One of the following holds:
+      \begin{enumproof}
+      \item (restrict,cca,seta2b) $w_1 = ((\perm',\lin),\baddr,\eaddr,\aaddr')$,
+        $w_1' = ((\perm',\lin),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) =
+        ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi_T(r_1) =
+        ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k =
+        1$
+      \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$,
+        $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) =
+        \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi_T(r_1) =
+        ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$,
+        $k = 1$
+      \item (cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$,
+        $w_1' = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $\Phi_S(r_1) =
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$ and $\Phi_T(r_1) =
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $k = 1$
+      \item (move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S(r_1)$, and
+        $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg \isLinear{w_1}
+        \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$) and $k = 2$.
+      \item (cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' =
+        \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) =
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k=1$
+      \item (split) $\Phi_T(r_3) = \Phi_S(r_3) =
+        ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$,
+        $w_1 = w_1' = ((\perm,\lin),\baddr,n,\aaddr)$, $w_2 = w_2' =
+        ((\perm,\lin),n+1,\eaddr,\aaddr)$, $k=2$.
+      \item (split) $\Phi_T(r_3) = \Phi_S(r_3) =
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \leq n$, $n <
+        \sigma_\eaddr$, $w_1 = w_1' = \seal{\sigma_\baddr,n,\sigma}$, $w_2 =
+        w_2' = \seal{n+1,\sigma_\eaddr,\sigma}$, $k=2$.
+      \item (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$,
+        $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$,
+        $n < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' =
+        ((\perm,\linear),\baddr,n,\aaddr)$, $w_2 =
+        \stkptr{\perm,n+1,\eaddr,\aaddr}$, $w_2' =
+        ((\perm,\linear),n+1,\eaddr,\aaddr)$, $k=2$.
+      \item (splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
+        $\Phi_S(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_}$, and $\Phi_T(r_3) =
+        ((\perm,\linear),\eaddr_2+1,\eaddr_3,\aaddr_3)$, $\Phi_S(r_3) =
+        \stkptr{\perm,\eaddr_2+1,\eaddr_3,\aaddr_3}$, and $\baddr_2 \leq
+        \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
+        $w_1=((\perm,\linear),\baddr_2,\eaddr_3,\aaddr_3)$, $w_1'=
+        \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_3}$, and $w_2 = w_2' = w_3 = w_3'
+        = 0$, and $k=3$
+      \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
+        ((\perm,\lin),\baddr_2,\eaddr_2,\_)$, and $\Phi_T(r_3) = \Phi_S(r_3) =
+        ((\perm,\lin),\eaddr_2+1,\eaddr_3,\aaddr_3)$, and $\baddr_2 \leq
+        \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
+        $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and $\lin =
+        \linear \Rightarrow w_2 = w_2' = w_3 = w_3' = 0$, and $\lin = \normal
+        \Rightarrow (w_2 = w_2' = \Phi_T(r_2) \text{ and } w_3 = w_3' =
+        \Phi_T(r_3))$, and $k=3$
+      \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
+        \seal{\sigma_{\baddr,2},\sigma_{\eaddr,2},\_}$, and $\Phi_T(r_3) =
+        \Phi_S(r_3) =
+        \seal{\sigma_{\eaddr,2}+1,\sigma_{\eaddr,3},\sigma_{\aaddr,3}}$, and
+        $\sigma_{\baddr,2} \leq \sigma_{\eaddr,2}$, and $\sigma_{\eaddr,2} + 1
+        \leq \sigma_{\eaddr,3}$and $w_1 = w_1' = \seal{\sigma_{\baddr,2},
+          \sigma_{\eaddr,3}, \sigma}$
+      \end{enumproof}
+    \end{itemize}
+  \item All of the following hold: (includes memory-manipulation cases: store,
+    load, that can be handled mostly uniformly)
     \begin{itemize}
-    \item (restrict,cca,seta2b) $w_1 = ((\perm',\lin),\baddr,\eaddr,\aaddr')$, $w_1' = ((\perm',\lin),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi_T(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k = 1$
-    \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$, $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$, $k = 1$
-    \item (cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $w_1' = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $\Phi_S(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$ and $\Phi_T(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $k = 1$
-    \item (move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S(r_1)$, and
-      $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg
-      \isLinear{w_1} \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$)
-      and $k = 2$.
-    \item (cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' = \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k=1$
-    \item (split) $\Phi_T(r_3) = \Phi_S(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$, $w_1 = w_1' = ((\perm,\lin),\baddr,n,\aaddr)$, $w_2 = w_2' = ((\perm,\lin),n+1,\eaddr,\aaddr)$, $k=2$.
-    \item (split) $\Phi_T(r_3) = \Phi_S(r_3) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \leq n$, $n < \sigma_\eaddr$, $w_1 = w_1' = \seal{\sigma_\baddr,n,\sigma}$, $w_2 = w_2' = \seal{n+1,\sigma_\eaddr,\sigma}$, $k=2$.
-    \item (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' = ((\perm,\linear),\baddr,n,\aaddr)$, $w_2 = \stkptr{\perm,n+1,\eaddr,\aaddr}$, $w_2' = ((\perm,\linear),n+1,\eaddr,\aaddr)$, $k=2$.
-    \item (splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
-      $\Phi_S(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_}$, and
-      $\Phi_T(r_3) = ((\perm,\linear),\eaddr_2+1,\eaddr_3,\aaddr_3)$,
-      $\Phi_S(r_3) = \stkptr{\perm,\eaddr_2+1,\eaddr_3,\aaddr_3}$, and
-      $\baddr_2 \leq \eaddr_2$, and
-      $\eaddr_2+1 \leq \eaddr_3$, and
-      $w_1=((\perm,\linear),\baddr_2,\eaddr_3,\aaddr_3)$, $w_1'=
-      \stkptr{\perm,\baddr_2,\eaddr_3,\aaddr_3}$, and
-      $w_2 = w_2' = w_3 = w_3' = 0$, and
-      $k=3$
-    \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
-      ((\perm,\lin),\baddr_2,\eaddr_2,\_)$, and
-      $\Phi_T(r_3) = \Phi_S(r_3) = ((\perm,\lin),\eaddr_2+1,\eaddr_3,\aaddr_3)$,
-      and
-      $\baddr_2 \leq \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
-      $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and
-      $\lin = \linear \Rightarrow w_2 = w_2' = w_3 = w_3' = 0$, and
-      $\lin = \normal \Rightarrow (w_2 = w_2' = \Phi_T(r_2) \text{ and } w_3 =
-      w_3' = \Phi_T(r_3))$, and $k=3$
-    \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) = \seal{\sigma_{\baddr,2},\sigma_{\eaddr,2},\_}$, and $\Phi_T(r_3) = \Phi_S(r_3) = \seal{\sigma_{\eaddr,2}+1,\sigma_{\eaddr,3},\sigma_{\aaddr,3}}$, and $\sigma_{\baddr,2} \leq \sigma_{\eaddr,2}$, and $\sigma_{\eaddr,2} + 1 \leq \sigma_{\eaddr,3}$and $w_1 = w_1' = \seal{\sigma_{\baddr,2}, \sigma_{\eaddr,3}, \sigma}$
-      \end{itemize}
-  \end{itemize}
-\item All of the following hold: (includes memory-manipulation cases: store, load, that can be handled mostly uniformly)
-  \begin{itemize}
-  \item $\Phi_S \step \Phi_S'$
+    \item $\Phi_S \step[\gc] \Phi_S'$
     \item $\Phi_T \step \Phi_T'$
@@
     \item One of the following hold:
-    \begin{itemize}
+      \begin{enumproof}
       \item (store) $w_1 = w_1' = \Phi_S(r_1) = \Phi_T(r_1) =
@@
         $\neg \isLinear{w_1} \Rightarrow w = w_1 \text{ and } w' = w_1'$
+      \end{enumproof}
     \end{itemize}
-  \end{itemize}
-\item All of the following hold: (includes memory-manipulation cases: store, load, that can be handled mostly uniformly (stack))
+  \item All of the following hold: (includes memory-manipulation cases: store,
+    load, that can be handled mostly uniformly (stack))
     \begin{itemize}
-  \item $\Phi_S \step \Phi_S'$
+    \item $\Phi_S \step[\gc]\Phi_S'$
     \item $\Phi_T \step \Phi_T'$
@@
     \item One of the following hold:
-    \begin{itemize}
-    \item (store) $w_1 = \Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $w_1' =  \Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \writeAllowed{}$,
+      \begin{enumproof}
+      \item (store) $w_1 = \Phi_T(r_1) =
+        ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $w_1' = \Phi_S(r_1) =
+        \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \writeAllowed{}$,
         and
@@
       \item (load) $ w_2' = \Phi_T(r_2) =
-      ((\perm,\linear),\baddr,\eaddr,\aaddr)$, and $w_2 = \Phi_S(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \readAllowed{}$,
+        ((\perm,\linear),\baddr,\eaddr,\aaddr)$, and $w_2 = \Phi_S(r_2) =
+        \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \readAllowed{}$,
         $\withinBounds{((\perm,\lin),\baddr,\eaddr,\aaddr)}$, and $\aaddr \in
@@
         $\neg \isLinear{w_1} \Rightarrow w = w_1 \text{ and } w' = w_1'$
+      \end{enumproof}
     \end{itemize}
-  \end{itemize}
-
-\item All of the following hold: (includes control-flow manipulation cases: jmp, jnz, xjmp, that can be handled mostly uniformly)
+  \item All of the following hold: (includes control-flow manipulation cases:
+    jmp, jnz, xjmp, that can be handled mostly uniformly)
     \begin{itemize}
-  \item $\Phi_S \step \Phi_S'$
+    \item $\Phi_S \step[\gc] \Phi_S'$
     \item $\Phi_T \step \Phi_T'$
     \item One of the following holds
-    \begin{itemize}
-    \item(jnz) $\Phi_S' = \updPcAddr{\Phi_S}$ and $\Phi_T' = \updPcAddr{\Phi_T}$
-    \item(jmp,jnz) $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and $\Phi_T' =
-      \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and 
-      $\isLinear{\Phi_S(r_1)} \Rightarrow w_1 = w_1' = 0$ and
-      $\neg \isLinear{\Phi_S(r_1)} \Rightarrow (w_1 = \Phi_S(r_1) \text{ and }
-      w_1' = \Phi_T(r_1))$
+      \begin{enumproof}
+      \item(jnz) $\Phi_S' = \updPcAddr{\Phi_S}$ and $\Phi_T' =
+        \updPcAddr{\Phi_T}$
+      \item(jmp,jnz) $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and
+        $\Phi_T' = \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and
+        $\isLinear{\Phi_S(r_1)} \Rightarrow w_1 = w_1' = 0$ and $\neg
+        \isLinear{\Phi_S(r_1)} \Rightarrow (w_1 = \Phi_S(r_1) \text{ and } w_1'
+        = \Phi_T(r_1))$
       \item(xjmp) $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) =
@@
         $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
-        $\nonExec{c_\data}$ and $\Phi_T'(\pcreg,\rdata) = c_1',c_2'$ and
-      one of the following hold:
-      \begin{itemize}
+        $\nonExec{c_\data}$ and $\Phi_T'(\pcreg,\rdata) = c_1',c_2'$ and one of
+        the following hold:
+        \begin{enumproof}
         \item $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$ and
@@
           \stkptr{\rw,\stkb,\eaddr_\stk,\_}$ and $\Phi_S.\stk = \stk_\var{frame}
-        :: (\aaddr,\ms_{\stk,\priv})$ and
-        $\dom(\ms_{\stk,\priv} = [\eaddr_\stk+1,\eaddr_{\stk,\priv}])$ and
-        $\Phi_S'.\ms_\stk = \ms_{\stk,\priv} \uplus \Phi_S.\ms_\stk$ and
-        $\Phi_S'.\stk = \stk$ and
+          :: (\aaddr,\ms_{\stk,\priv})$ and $\dom(\ms_{\stk,\priv} =
+          [\eaddr_\stk+1,\eaddr_{\stk,\priv}])$ and $\Phi_S'.\ms_\stk =
+          \ms_{\stk,\priv} \uplus \Phi_S.\ms_\stk$ and $\Phi_S'.\stk = \stk$ and
           \[
@@
           \]
+        \end{enumproof}
+      \end{enumproof}
     \end{itemize}
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
+  \end{enumproof}
+\lau{TODO: Argue that the above is indeed the case.}
+
 \end{proof}

### 2018-05-24 fully-abstract-comp/technical-report.tex
Fix a lemma statement and first step of the proof.
Updated FTLR proof refactoring so it is easier to refer to specific cases.

--- 2018-05-23 fully-abstract-comp/technical-report.tex
@@
 \usepackage{url}
-\usepackage{hyperref}
+
+\usepackage{enumitem}
 
@@
 
-A component needs access to the following:
+A program needs access to the following:
 \begin{description}
@@
     \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
-    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}}
   }{
@@
     \msdata[3] = \msdata[1] \uplus \msdata[2] \\
-    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
-    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
+    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}&
+    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}\\
     \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
@@
 \end{mathpar}
-
-\begin{lemma}[Linking preserves component well-formedness]
-  \label{lem:linking-well-formedness}
-  If $\vdash \var{comp}_1$ and $\vdash \var{comp}_2$, then $\vdash \var{comp}_1 \bowtie \var{comp}_2$.
-\end{lemma}
-\begin{proof}
-  TODO
-\end{proof}
 
@@
 
+\begin{lemma}[purePart is idempotent]
+  \label{lem:purePart-idempotemt}
+  For all $W$, we have that $\purePart{W} = \purePart{}(\purePart{W})$ and $\purePart{W} = \purePart{W} \oplus \purePart{W}$.
+\end{lemma}
+
 \begin{lemma}[purePart is duplicable]
@@
 \end{lemma}
-\begin{proof}
-  Follows from the definition of $\purePart{}$ and $\oplus$.
-\end{proof}
-
-\begin{lemma}[purePart is idempotent]
-  \label{lem:purePart-idempotent}
-  For all $W$, we have that $\purePart{W} = \purePart{}(\purePart{W})$ and $\purePart{W} = \purePart{W} \oplus \purePart{W}$.
-\end{lemma}
-\begin{proof}
-  The first part follows easily from the definition.
-  The second statement follows from the first, together with Lemma~\ref{lem:purePart-duplicable}.
-\end{proof}
 
@@
 \end{lemma}
-\begin{proof}
-  Follows easily from the definition of $\purePart{}$ and $\future$.
-\end{proof}
 
@@
 \end{lemma}
-\begin{proof}
-  Follows easily from the definitions.
-\end{proof}
 
@@
 \end{lemma}
-\begin{proof}
-  Consequence of Lemmas~\ref{lem:purePart-duplicable}
-  and~\ref{lem:oplus-future}.
-\end{proof}
 
@@
       \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; W \text{ for all $n' < n$}\wedge\\
+      \quad \stpair[.]{\vsc_S}{\vsc_T} \in H_\sigma \; \sigma \; W\wedge \\
       \quad (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
       \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad\forall W' \future W, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
       \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \stpair[.]{\vsc_S'}{\vsc'_T} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
@@
         \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
-        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
-        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \readCond{\lin,W} \wedge\\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{(\baddr,\eaddr)} \in \writeCond{\lin,W} \wedge\\
         % we are excluding rwx pointers.
@@
         \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
-                                  \npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}\endarray
+                                  \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}\endarray
       \end{array}
@@
       \perm \sqsubseteq \rx \wedge \\
-      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,(\ta,\stkb)]{\normal,W}
+      \npair{(\baddr,\eaddr)} \in \xReadCond[\square,(\ta,\stkb)]{\normal,W}
     \end{array}
@@
 \[
-  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
+  \readCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
     \begin{array}{l}
@@
       \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
       \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
@@
 \[
-  \xReadCond{\lin,W} = \left\{ \npair{A} \middle| 
+  \xReadCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr})) \middle| 
     \begin{array}{l}
@@
       \quad \exists M : S \fun \powerset{\MemSeg} \\
-      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
+      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr] \wedge\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
@@
 \[
-  \writeCond{\lin,W} = \left\{ \npair{A} \middle| 
+  \writeCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)}) \middle| 
     \begin{array}{l}
@@
       \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge\\
       \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
@@
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-    \dom(\code) \subseteq \ta \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset) \wedge \\
+    \dom(\code) \subseteq \ta \Rightarrow \sigrets = \emptyset \wedge \\
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
@@
 \left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
-      \begin{array}{l}
-        \dom(\code) \subseteq \ta \tand\\
+      \\
+      \begin{array}{l}
         \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
@@
 \left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
-      \begin{array}{l}
-        \dom(\code) \subseteq \ta \tand\\
+      \\
+      \begin{array}{l}
         \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
@@
 
-\subsection{World specifies component}
+\subsection{Fundamental Theorem of Logical Relations}
 
@@
   \item $\mscode = \biguplus_i \code_i$
-  \item For all $i$, one of the following holds:
-    \begin{itemize}
-    \item $\dom(\code_i) \subseteq \ta$
-    \item $\dom(\code_i) \mathrel{\#} \ta$
-    \end{itemize}
-  \item $\sigrets = \biguplus_i \sigrets_i$
-  \item $\sigcloss = \biguplus_i \sigcloss_i$.
+  \item One of the following holds:
+    \begin{itemize}
+    \item $\dom(\mscode) \subseteq \ta$ and $\sigrets = \biguplus_i \sigrets_i$ and $\sigcloss = \biguplus_i \sigcloss_i$.
+    \item $\dom(\mscode) \mathrel{\#} \ta$ and $\sigrets = \emptyset$ and $\sigcloss = \biguplus_i \sigrets_i \uplus \biguplus_i \sigcloss_i$.
+    \end{itemize}
   \end{itemize}
@@
   \begin{itemize}
-  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\lin}$ with $\lin = \pure$ or $\lin = \spatialo$
+  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\pure}$
   \item $\dom(\msdata) = \biguplus_i A_i$
@@
 
-\begin{lemma}[World specifies linked component if it specifies each separately]
-  \label{lem:world-specifies-linked}
-  If
-  \begin{itemize}
-  \item $W \vdash_\gc \var{comp}_0$
-  \item $W \vdash_\gc \var{comp}_0'$
-  \end{itemize}
-  Then $W \vdash \var{comp}_{0} \bowtie \var{comp}_{0}'$
-\end{lemma}
-\begin{proof}
-  Follows easily from the definitions.
-\end{proof}
-
-\begin{lemma}[World specifies component pure]
-  \label{lem:world-specifies-comp-pure}
-  If $W \vdash_\gc \var{comp}_0$ then $\purePart{W} \vdash \var{comp}_{0}$.
-\end{lemma}
-\begin{proof}
-  Follows easily from the definitions.
-\end{proof}
-
-\subsection{Fundamental Theorem of Logical Relations}
 \begin{theorem}[FTLR]
@@
   \begin{itemize}
-  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
-  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}$
+  \item $\trust = \trusted \vee \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
+  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
   \end{itemize}
@@
       W \vdash \var{comp} \tand\\
-      \text{For all } W' \future W \ldotp \text{If } \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+      \text{For all } W' \future W \ldotp \text{If } \overline{\npair{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'})\\
       \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
@@
      \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
-     \{c_{\mathrm{main},c},c_{\mathrm{main},d}\} \subseteq \overline{c_{\mathrm{export}}}
+      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+     \{(c_{\mathrm{main},c}, c_{\mathrm{main},c}),(c_{\mathrm{main},d}, c_{\mathrm{main},d})\} \subseteq (\overline{c_{\mathrm{export}}},\overline{c_{\mathrm{export}}})
        \end{multlined}
@@
   \begin{itemize}
-  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$
-  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$
+  \item $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$
+  \item $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$
   \end{itemize}
@@
   \begin{itemize}
-  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W)$.
+  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W_1\oplus W_2)$.
   \end{itemize}
 \end{lemma}
-\begin{proof}
-  First consider the case where one of the two pairs of components have a pair of ``main'' capabilities.
-  By the definition of $\lrcomp$, these capabilities have to be part of that components' export list.
-  It is then easy to see that the other component cannot have ``main'' capabilities (otherwise linking is not defined) and it is sufficient to prove the result for the underlying components (without the ``main'' capabilities).
-  Therefore, we can restrict ourselves to the case where both pairs of components are of the form $\var{comp}_0$, i.e. no ``main'' capabilities.
-
-  By Lemma~\ref{lem:world-specifies-linked}, we have that $W \vdash \var{comp}_1\bowtie \var{comp}_2$.
-
-  By definition of $\bowtie$, we have that the following hold:
-  \begin{align*}
-    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
-    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
-    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
-    \mscode[3] = \mscode[1] \uplus \mscode[2] &
-    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
-    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
-    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
-    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
-    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
-  \end{align*}
-
-  Now take $W' \future W$ and assume that $\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
-  Then it remains to show that $\overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$.
-  We do this by complete induction on $n$, so that we can assume that $\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
-
-  For each $$\npair{(c_{\mathrm{export}},c_{\mathrm{export}})} \in \overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}}$$, we have that $\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}$ is either in $\overline{\npair{(c_{\mathrm{export},1},c_{\mathrm{export},1})}}$ or $\overline{\npair{(c_{\mathrm{export},2},c_{\mathrm{export},2})}}$.
-  By unfolding $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$, it suffices to prove that $$\overline{\npair[n']{(c_{\mathrm{import},1},c_{\mathrm{import},1})}} \in \lrv(\purePart{W'})$$ and $$\overline{\npair[n']{(c_{\mathrm{import},2},c_{\mathrm{import},2})}} \in \lrv(\purePart{W'})$$
-  for all $n' < n$.
-
-  Simple set algebra tells us that 
-  $$\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}} \subseteq \overline{c_{\mathrm{export},3}} \cup \overline{c_{\mathrm{import},3}}$$
-  and for all $n' < n$, we know that $$\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$$
-  and
-  $$\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})\text{.}$$
-\end{proof}
 
@@
 
-\begin{lemma}[FTLR for component data-values]
-  If
-  \begin{equation*}
-    \dom(\mscode), A_{\mathrm{own}}, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
-  \end{equation*}
-  and
-  \begin{itemize}
-  \item $\npair{A_{\mathrm{non-linear}}} \in \readCond{\normal,W}$
-  \item $\npair{A_{\mathrm{non-linear}}} \in \writeCond{\normal,W}$
-  \item $\npair{A_{\mathrm{own}}} \in \readCond{\linear,W}$
-  \item $\npair{A_{\mathrm{own}}} \in \writeCond{\linear,W}$
-  \item $W(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
-  \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
-  \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(W)$ for all $n' < n$
-  \item $w \not\in \overline{c_{\mathrm{import}}}$
-  \end{itemize}
-
-  then $\npair{(w,w)} \in \lrv(W)$
-\end{lemma}
-\begin{proof}
-  By induction on the judgement 
-  \begin{equation*}
-    \dom(\mscode), A_{\mathrm{own}}, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
-  \end{equation*}
-  We have the following cases:
-  \begin{itemize}
-  \item $w = z$: result follows trivially
-  \item $w = c$, $c \in \overline{c_{\mathrm{import}}}$: contradicts assumption.
-  \item $w = ((\perm,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq A_{\mathrm{code}}$, $\perm \sqsubseteq \rx$:
-    TODO
-  \item $w = ((\perm,\lin),\baddr,\eaddr,\aaddr)$,
-    $\permbnf \sqsubseteq \rw$,
-    $\lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}}$ and
-    $\lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}$:
-    TODO
-
-  \item $w = \sealed{\sigma,\vsc}$ and
-    \begin{equation*}
-      A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc
-    \end{equation*}
-    and $\sigma \in \sigcloss$:
-
-    By definition of $\lrv$ and by choosing island $r = r_{\mathrm{code}}$, it suffices to prove that
-    \begin{itemize}
-    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; W$ for all $n' < n$:
-
-     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that
-     \begin{align*}
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc)} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\hat{W}))
-     \end{align*}
-
-     TODO
-
-     this follows
-     by induction and Lemma~\ref{lem:non-expansive} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption, we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
-   \item $(\isLinear{\vsc} \text{ iff } \isLinear{\vsc})$: trivially fine
-
-   \item If $\isLinear{\vsc}$, then 
-     for all $W' \future W$, $W_o$, $n' < n$ and $\npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that 
-     \begin{equation*}
-       \npair[n']{(\vsc,\vsc_S',\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o)) \wedge \\
-     \end{equation*}
-     TODO
-   \item 
-       If $\nonLinear{\src{\vsc_S}}$ then for all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
-       \begin{equation*}
-         \npair[n']{(\vsc,\src{\vsc_S'},\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
-       \end{equation*}
-       TODO
-    \end{itemize}
-  \end{itemize}
-\end{proof}
-
 \begin{lemma}[FTLR for components]
@@
   \item $\gc = (\ta,\stkb)$
-  \item $\var{comp}$ is a well-formed component, i.e. $\vdash \var{comp}$
   \item One of the following holds:
     \begin{itemize}
-    \item $\var{comp}$ is a reasonable component (see
-      Section~\ref{sec:reas-comp}) and $\dom(\var{comp}.\mscode) \subseteq \ta$
-    \item $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
+    \item $\trust = \trusted$, $\var{comp}$ is a reasonable component (see
+      Section~\ref{sec:full-abstraction}) and $\dom(\comp.\mscode) \subseteq \ta$
+    \item $\trust = \untrusted$ and $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
     \end{itemize}
@@
 \end{lemma}
-\begin{proof}
-  If $\var{comp}$ is of the form $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$, then we have (by definition of component well-formedness) that $c_{\mathrm{main},c}, c_{\mathrm{main},d} \subseteq \var{comp}_0.\overline{c_{\mathrm{export}}}$, as required by $\lrcomp(W)$.
-  Hence, we can restrict our attention to components of the form $\var{comp}_0$.
-  Take
-  \begin{equation*}
-    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
-  \end{equation*}
-
-  We know that $W \vdash \var{comp}$ by assumption, so it remains to prove that 
-  for all  $W' \future W$ with 
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
-  \end{equation*}
-  we have that
-  \begin{equation*}
-    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
-  \end{equation*}
-
-  TODO
-\end{proof}
 Note: the trusted case of the above lemma can be considered as a compiler correctness result.
@@
 \lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}
-
-\subsection{Reasonable components}
-\label{sec:reas-comp}
 \begin{definition}
@@
 
-%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}
+%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
 \begin{definition}[Never store link seals and code capability]
@@
 \begin{definition}
-  A component $\var{comp})$ with $\dom(\var{comp}.\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
-  \begin{itemize}
-  \item $\plug{\context}{\var{comp}}$ checks $\stkb$ before calls in $A$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-  \item $\plug{\context}{\var{comp}}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\plug{\context}{\var{comp}}$ handle trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-  \item $\plug{\context}{\var{comp}}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+  A component $\program = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)$ with $\dom(\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
+  \begin{itemize}
+  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $A$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
+  \item $\plug{\context}{\program}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
+  \item $\plug{\context}{\program}$ handle trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
   \end{itemize}
@@
 
-\subsection{Full abstraction: precise formulation}
-\label{sec:full-abstr-prec}
-
 \begin{definition}[Source language contextual equivalence]
-  In the source language, we define that $\src{\var{comp}_1 \sconeq \var{comp}_2}$ iff
+  In the source language, we define that $\src{\program_1 \sconeq \program_2}$ iff
   \begin{equation*}
-    \forall \src{\context} \ldotp \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
+    \forall \src{\context} \ldotp \src{\plug{\context}{\program_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\program_2} \sterm{\ta[,2],\stkb_2}}
   \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$
+  with $\src{\ta[,i]} = \src{\dom(\program_i.\mscode)}$
 \end{definition}
@@
 \begin{definition}[Target language contextual equivalence]
-  In the target language, we define that $\trg{\var{comp}_1 \tconeq \var{comp}_2}$ iff
+  In the target language, we define that $\trg{\program_1 \tconeq \program_2}$ iff
   \begin{equation*}
-    \forall \trg{\context} \ldotp \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
+    \forall \trg{\context} \ldotp \trg{\plug{\context}{\program_1} \term} \Leftrightarrow \trg{\plug{\context}{\program_2} \term}
   \end{equation*}
@@
   \label{thm:full-abstraction}
-  For reasonable components $\var{comp}_1$ and $\var{comp}_2$, we have
+  For reasonable components $\program_1$ and $\program_2$, we have
   \begin{gather*}
-    \src{\var{comp}_1} \sconeq \src{\var{comp}_2}\\
+    \src{\program_1} \sconeq \src{\program_2}\\
     \Updownarrow\\
-    \comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}
+    \comp{\src{\program_1}} \tconeq \comp{\src{\program_2}}
   \end{gather*}

### 2018-05-24 fully-abstract-comp/technical-report.tex
Downwards closure lemmas about the conditions on source configurations

@@
 
-\section{Proofs and lemmas}
+\section{Lemmas}
+\subsection{Well-behaved configurations}
+\begin{lemma}
+  \label{lem:check-stack-addr-before-call-dc}
+For $n' \leq n$, if a configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps, then $\src{\Phi'}$ is said to check $\stkb$ before calls in $\ta$ for $n'$ steps
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:use-return-seals-call-dc}
+  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:handle-trusted-seals-prop-dc}
+  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:never-store-seal-code-cap-dc}
+  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\section{Proofs}
 \input{ftlr-proof-cases}

### 2018-05-24 fully-abstract-comp/ftlr-proof-cases.tex

@@
   \begin{itemize}
-  \item If $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$, then $\npair[n']{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$.
-  \item If $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$, then $\npair[n']{(\baddr,\eaddr)} \in \readCond{\lin,W}$.
-  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
-  \item If $\npair{(\reg_S,\reg_T)} \in \lrr(W)$, then $\npair[n']{(\reg_S,\reg_T)} \in \lrr(W)$.
+  \item If $\npair{A} \in \readCond{\lin,W}$, then $\npair[n']{A} \in \readCond{\lin,W}$.
+  \item If $\npair{A} \in \stackReadCond{\lin,W}$, then $\npair[n']{A} \in \stackReadCond{\lin,W}$.
+  \item If $\npair{A} \in \writeCond{\lin,W}$, then $\npair[n']{A} \in \writeCond{\lin,W}$.
+  \item If $\npair{A} \in \stackWriteCond{\lin,W}$, then $\npair[n']{A} \in \stackWriteCond{\lin,W}$.
+  \item If $\npair{A} \in \execCond{\lin,W}$, then $\npair[n']{A} \in \execCond{\lin,W}$.
+  \item If $\npair{A} \in \xReadCond{\lin,W}$, then $\npair[n']{A} \in \xReadCond{\lin,W}$.
+  \item If $\npair{(w,w')} \in \lrvg{\trust}{W}$, then $\npair[n']{(w,w')} \in \lrvg{\trust}{W}$.
+  \item If $\npair{(w,w')} \in \lrrg{\trust}(W)$, then $\npair[n']{(w,w')} \in \lrrg{\trust}(W)$.
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}(W)$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}(W)$.
   \end{itemize}
@@
   \item ...
-  \end{itemize}
 \end{lemma}
@@
   From this, it follows that $\memSatFStack{\ms_\stk\update{a}{0},\ms_\var{T,heap}\update{a}{0}}{W_r' \oplus W_\var{free\_stack}'}$ and finally $\memSat{(\ms_S,\stk,\ms_\stk\update{a}{0},\ms_T\update{a}{0})}{W_M'}$.
-\end{proof}
-
-\begin{lemma}[purePart is duplicable]
-  \label{lem:purePart-duplicable}
- For all $W$, we have that $W = W \oplus \purePart{W}$.
-\end{lemma}
-\begin{proof}
-  Follows from the definition of $\purePart{}$ and $\oplus$.
-\end{proof}
-
-\begin{lemma}[purePart is idempotent]
-  For all $W$, $\mathit{purePart}( \purePart{W} ) = \purePart{W}$ and $\purePart{W} = \purePart{W} \oplus \purePart{W}$.
-  \label{lem:purePart-idempotent}
-\end{lemma}
-\begin{proof}
-  The first part follows easily from the definition.
-  The second statement follows from the first, together with Lemma~\ref{lem:purePart-duplicable}.
-\end{proof}
-
-\begin{lemma}[purePart is monotone]
-  \label{lem:purePart-mono}
- For all $W'\future W$, we have that $\purePart{W'} \future \purePart{W}$.
-\end{lemma}
-\begin{proof}
-  Follows easily from the definition of $\purePart{}$ and $\future$.
-\end{proof}
-
-\begin{lemma}[Increasing authority is the future]
-  \label{lem:oplus-future}
-  For all $W$, we have that:
-  $W_1 \oplus W_2 \future W_1$
-\end{lemma}
-\begin{proof}
-  Follows easily from the definitions.
-\end{proof}
-
-\begin{lemma}[Purity is a thing of the past]
-  \label{lem:world-fut-purePart}
-  For all $W$, we have that $W \future \purePart{W}$.
-\end{lemma}
-\begin{proof}
-  Consequence of Lemmas~\ref{lem:purePart-duplicable}
-  and~\ref{lem:oplus-future}.
 \end{proof}

### 2018-05-24 fully-abstract-comp/technical-report.tex

--- 2018-05-23 fully-abstract-comp/technical-report.tex
@@
 \usepackage{hyperref}
+
+\usepackage{enumitem}
 
@@
 \begin{lemma}[FTLR for component code-values]
-  If $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} w$
-  and
+  If
   \begin{itemize}
@@
   \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
+  \item $w = ((\perm,\rx),\baddr,\eaddr,\aaddr)$
   \end{itemize}
@@
 \begin{proof}
-  By induction on the judgement $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} w$.
-
+  We distinguish two cases:
+  \begin{itemize}
+  \item $\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$:
     TODO
+  \item $\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$:
+    TODO
+  \end{itemize}
 \end{proof}

### 2018-05-24 fully-abstract-comp/ftlr-proof-cases.tex
The new cases to the FTLR proof.

@@
 
+By the above observation, we know that $\Phi_S \step[\gc] \Phi_S'$ and $\Phi_T \step
+\Phi_T'$ for some $\Phi_S'$ and $\Phi_T'$. According to
+Lemma~\ref{lem:lro-anti-red-gen}, it suffices to show:
+\[
+  \npair[n'-1]{(\Phi_S',\Phi_T')} \in \lro
+\]
+Consider each of the possible cases:
+
+In case \ref{case:ftlr:failed}, both executions go to failed. In this case, we
+either assume that the target configuration terminates or that the source
+configuration terminates, but in either case it leads to a contradiction with
+the failing configuration.
+
+In case \ref{case:ftlr:get-and-arith-op}, we use the induction hypothesis to get
+\[
+  \npair[n'-1]{(((\perm,\lin),\baddr,\eaddr,\aaddr+1),((\perm,\lin),\baddr,\eaddr,\aaddr+1))} \in \lre(W)
+\]
+This is possible by using the initial assumptions along with Lemma \ref{lem:check-stack-addr-before-call-dc}, \ref{lem:use-return-seals-call-dc}, \ref{lem:handle-trusted-seals-prop-dc}, \ref{lem:never-store-seal-code-cap-dc}, and \ref{lem:downwards-closed}.
+By Lemma~\ref{lem:downwards-closed}, we know $\memSat[n'-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
+By Lemma~\ref{lem:downwards-closed} and the fact that integers are always safe,
+we know $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$.
+This allows us to conclude
+\[
+  \npair[n'-1]{(\Phi_S',\Phi_T')} \in \lro
+\]
+as desired.
+
 \end{proof}

### 2018-05-24 fully-abstract-comp/technical-report.tex
Some work on FTLR for components proof.

@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,((\perm,\rx),\baddr,\eaddr,\aaddr)}
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
   }
@@
 
-\begin{lemma}[FTLR for component code-values]
+\begin{lemma}[Code region for untrusted components stronger than standard safe memory region]
+  \label{lem:codereg-untrusted-stdreg}
   If
   \begin{itemize}
-  \item $W(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item $\dom(\mscode) \mathrel{\#} \ta$
+  \end{itemize}
+  then $\codereg{\sigrets,\sigcloss,\mscode,\gc} \nsubeq \stdreg{R(r),\gc}{\pur}$ 
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}[Code memory for untrusted components safely readable]
+  \label{lem:code-memory-untrusted-readable}
+  If
+  \begin{itemize}
+  \item $\dom(\mscode) \mathrel{\#} \ta$
+  \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item $A \subseteq \dom(\mscode)$,
+  \end{itemize}
+  we have that $\npair{A} \in \readCond{\normal,W}$.
+\end{lemma}
+\begin{proof}
+  Follows by defintion of $\readCond$ from Lemma~\ref{lem:codereg-untrusted-stdreg}.
+\end{proof}
+
+\begin{lemma}[FTLR for component code capabilities]
+  \label{lem:ftlr-component-code-ptrs}
+  If
+  \begin{itemize}
+  \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
   \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
-  \item $w = ((\perm,\rx),\baddr,\eaddr,\aaddr)$
+  \item $w = ((\rx,\normal),\baddr,\eaddr,\aaddr)$
+  \item $[\baddr,\eaddr]\subseteq \dom(\mscode)$
   \end{itemize}
@@
   \item $\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$:
-    TODO
+    By definition of $\lrv(W)$, it suffices to show that:
+    $\npair{[b,e]}\in\readCond{\normal,W}$, 
+    $\npair{[b,e]}\in\xReadCond{\normal,W}$ and 
+    $\npair{[b,e]}\in\execCond{\normal,W}$.
+
+    The first requirement follows from Lemma~\ref{lem:code-memory-untrusted-readable}.
+    The second follows from the definition of $\xReadCond$, together with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+    The third TODO \dominique{TODO: assume induction hypo of FTLR?}
   \item $\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$:
-    TODO
+    By definition of $\lrvg{\trusted}(W)$, it suffices to show that
+    $\npair{[b,e]}\in\xReadCond{\normal,W}$.
+
+    This follows immediately from the definition of $\xReadCond$, together with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
   \end{itemize}
@@
   \item $\npair{A_{\mathrm{own}}} \in \writeCond{\linear,W}$
-  \item $W(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
   \item $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$
@@
   \item $w = c$, $c \in \overline{c_{\mathrm{import}}}$: contradicts assumption.
-  \item $w = \sealed{\sigma,((\perm,\rx),\baddr,\eaddr,\aaddr)}$, $[\baddr,\eaddr]\subseteq A_{\mathrm{code}}$, $\sigma \in \sigcloss$:
-
+  \item $w = \sealed{\sigma,\vsc}$, $\vsc = ((\rx,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq \dom(\mscode)$, $\sigma \in \sigcloss$:
+
+    By definition of $\lrv$ and by choosing island $r = r_{\mathrm{code}}$, it suffices to prove that
+    \begin{itemize}
+    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
+
+     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that one of the following holds:
+     \begin{itemize}
+     \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrv \; \xi(\xi^{-1}(W))$
+     \item $(\dom(\code) \subseteq \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\xi^{-1}(W))$
+     \end{itemize}
+     Take $\trust = \untrusted$ iff $\dom(\mscode \mathrel{\#} \ta)$ and $\trust = \trusted$ iff $\dom(\mscode) \subseteq \ta$.
+     Since $\xi(\xi^{-1}(W)) = W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrvg{\trust}(W)$.
+
+     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs} that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(W)$, using Lemma~\ref{lem:downwards-closed}.
+
+   \item $(\isLinear{\vsc} \text{ iff } \isLinear{\vsc})$: trivially fine
+
+   \item 
+     For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+     \begin{equation*}
+       \npair[n']{(\vsc,\src{\vsc_S'},\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
+     \end{equation*}
+     \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
+   \end{itemize}
    TODO
@@
     \begin{equation*}
-      A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc
+      \dom(\mscode),A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc
     \end{equation*}
@@
 
-     This follows by induction and Lemma~\ref{lem:non-expansive} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
 

### 2018-05-24 fully-abstract-comp/ftlr-proof-cases.tex
reordered cases in FTLR proof.
Compilation issue fixed
Removed steps from well-behaved source configuration conditions.

@@
   \item ...
+  \end{itemize}
 \end{lemma}
@@
       \begin{itemize}           
-      \item $\src{\Phi}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-      \item $\src{\Phi}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-      \item $\src{\Phi}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-      \item $\src{\Phi}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
+      \item $\src{\Phi}$ checks $\stkb$ before calls in $\ta$ (Definition~\ref{def:check-stack-addr-before-call}).
+      \item $\src{\Phi}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ (Definition~\ref{def:use-return-seals-call}).
+      \item $\src{\Phi}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ (Definition~\ref{def:handle-trusted-seals-properly}).
+      \item $\src{\Phi}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ (Definition~\ref{def:never-store-seal-code-cap}).
       \end{itemize}
@@
   \item $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
-  \item All of the following hold:
-    \begin{itemize}
-    \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
-    \item $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$
-    \item $\exec{\Phi_S(\pcreg)}$
-    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in
-      \ints$.
-    \item TODO: other useful info
-    \end{itemize}
   \item \label{case:ftlr:get-and-arith-op} All of the following hold: (includes simple cases: gettype, geta, getb,
@@
     \item $z \in \ints$
+    \end{itemize}
+  \item All of the following hold:
+    \begin{itemize}
+    \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
+    \item $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$
+    \item $\exec{\Phi_S(\pcreg)}$
+    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in
+      \ints$.
+    \item TODO: other useful info
     \end{itemize}
@@
 \]
-This is possible by using the initial assumptions along with Lemma \ref{lem:check-stack-addr-before-call-dc}, \ref{lem:use-return-seals-call-dc}, \ref{lem:handle-trusted-seals-prop-dc}, \ref{lem:never-store-seal-code-cap-dc}, and \ref{lem:downwards-closed}.
+This is possible by using the initial assumptions along with Lemma \ref{lem:downwards-closed}.
 By Lemma~\ref{lem:downwards-closed}, we know $\memSat[n'-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.

### 2018-05-24 fully-abstract-comp/technical-report.tex
Removed steps from well-behaved source configuration conditions.

--- 2018-05-24 fully-abstract-comp/technical-report.tex
@@
   \label{def:check-stack-addr-before-call}
-  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
-  for all $i \leq n$ if
+  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ iff\\
+  for all $i$ if
   \[
@@
   \label{def:use-return-seals-call}
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
-  for all $i \leq n$, if
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
+  for all $i$, if
   \[
@@
   \label{def:handle-trusted-seals-properly}
-  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
-  for all $i \leq n$, if
+  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
+  for all $i$, if
   \[
@@
   \label{def:never-store-seal-code-cap}
-  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
-  for all $i \leq n$, if
+  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
+  for all $i$, if
   \[
@@
 \section{Lemmas}
-\subsection{Well-behaved configurations}
-\begin{lemma}
-  \label{lem:check-stack-addr-before-call-dc}
-For $n' \leq n$, if a configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps, then $\src{\Phi'}$ is said to check $\stkb$ before calls in $\ta$ for $n'$ steps
-\end{lemma}
-\begin{proof}
-  \lau{TODO}
-\end{proof}
-
-\begin{lemma}
-  \label{lem:use-return-seals-call-dc}
-  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
-\end{lemma}
-\begin{proof}
-  \lau{TODO}
-\end{proof}
-
-\begin{lemma}
-  \label{lem:handle-trusted-seals-prop-dc}
-  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
-\end{lemma}
-\begin{proof}
-  \lau{TODO}
-\end{proof}
-
-\begin{lemma}
-  \label{lem:never-store-seal-code-cap-dc}
-  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
-\end{lemma}
-\begin{proof}
-  \lau{TODO}
-\end{proof}
 

### 2018-05-24 fully-abstract-comp/ftlr-proof-cases.tex
added some labels, fixed reference.

@@
   \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T \step\failed$
-  \item $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
+  \item \label{case:ftlr:halted} $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
   \item \label{case:ftlr:get-and-arith-op} All of the following hold: (includes simple cases: gettype, geta, getb,
@@
     \end{itemize}
-  \item All of the following hold:
+  \item \label{case:ftlr:call} All of the following hold:
     \begin{itemize}
@@
 the failing configuration.
-
+\\\\
+In case \ref{case:ftlr:halted}, both source and target configuration halts in 0
+steps, so both directions of $\lro$ are trivially satisfied.
+\\\\
 In case \ref{case:ftlr:get-and-arith-op}, we use the induction hypothesis to get

### 2018-05-24 fully-abstract-comp/technical-report.tex
added some labels, fixed reference.

@@
 
-     This follows by induction and Lemma~\ref{lem:non-expansive} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
 

### 2018-05-24 fully-abstract-comp/technical-report.tex

--- 2018-05-24 fully-abstract-comp/technical-report.tex
@@
      Since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$ in both cases.
-
-     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
+     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
 

### 2018-05-24 fully-abstract-comp/technical-report.tex
Some more work on the FTLR for components.

@@
 
+\begin{lemma}[readCond shrinkable]
+  \label{lem:readCond-shrinkable}
+  If $\npair{A} \in \readCond{\lin,W}$  and $A' \subseteq A$, then
+  $\npair{A'} \in \readCond{\lin,W}$.
+\end{lemma}
+\begin{proof}
+  Follows easily from the definition.
+\end{proof}
+
+\begin{lemma}[readXCond shrinkable]
+  \label{lem:readXCond-shrinkable}
+  If $\npair{A} \in \xReadCond{\lin,W}$  and $A' \subseteq A$, then
+  $\npair{A'} \in \xReadCond{\lin,W}$.
+\end{lemma}
+\begin{proof}
+  Follows easily from the definition.
+\end{proof}
+
 \begin{lemma}[FTLR for component code capabilities]
@@
     By definition of $\lrv(W)$, it suffices to show that:
-    $\npair{[b,e]}\in\readCond{\normal,W}$, 
-    $\npair{[b,e]}\in\xReadCond{\normal,W}$ and 
-    $\npair{[b,e]}\in\execCond{\normal,W}$.
-
-    The first requirement follows from Lemma~\ref{lem:code-memory-untrusted-readable}.
-    The second follows from the definition of $\xReadCond$, together with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
-    The third TODO \dominique{TODO: assume induction hypo of FTLR?}
+    \begin{itemize}
+    \item $\npair{[b,e]}\in\readCond{\normal,W}$:
+      this follows from Lemma~\ref{lem:code-memory-untrusted-readable}.
+    \item $\npair{[b,e]}\in\xReadCond{\normal,W}$:
+      follows from the definition of $\xReadCond$, together with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+    \item $\npair{[b,e]}\in\execCond{\normal,W}$:
+      take $n' < n$, $W' \future \purePart{W}$, $\aaddr' \in [\baddr',\eaddr'] \subseteq [\baddr,\eaddr]$ and $w' = ((\rx,\normal),\baddr',\eaddr',\aaddr')$.
+      Then we need to show that $\npair[n']{(w',w')} \in \lre(W')$.
+      This now follows immediately from the (regular) FTLR (Theorem~\ref{thm:ftlr}), using the two above points, Lemma~\ref{lem:downwards-closed} and Lemmas~\ref{lem:readCond-shrinkable} and~\ref{lem:readXCond-shrinkable}.
+    \end{itemize}
+
   \item $\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$:
@@
      Since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$ in both cases.
-     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+
+     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
 
@@
   \label{def:check-stack-addr-before-call}
-  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ iff\\
-  for all $i$ if
+  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
+  for all $i \leq n$ if
   \[
@@
   \label{def:use-return-seals-call}
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
-  for all $i$, if
+  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  for all $i \leq n$, if
   \[
@@
   \label{def:handle-trusted-seals-properly}
-  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
-  for all $i$, if
+  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  for all $i \leq n$, if
   \[
@@
   \label{def:never-store-seal-code-cap}
-  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
-  for all $i$, if
+  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
+  for all $i \leq n$, if
   \[
@@
 \section{Lemmas}
+\subsection{Well-behaved configurations}
+\begin{lemma}
+  \label{lem:check-stack-addr-before-call-dc}
+For $n' \leq n$, if a configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps, then $\src{\Phi'}$ is said to check $\stkb$ before calls in $\ta$ for $n'$ steps
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:use-return-seals-call-dc}
+  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:handle-trusted-seals-prop-dc}
+  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  \label{lem:never-store-seal-code-cap-dc}
+  For $n' \leq n$, if a configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps, then $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n'$ steps.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
 

### 2018-05-29 fully-abstract-comp/ftlr-proof-cases.tex
Updated ftlr cases to let failing happen in a number of steps.
Added some more to the call case in the refactoring.
Add some more lemmas about splittability, shrinkability and purity of permission-based conditions and the value relation.

@@
 
-\begin{lemma}[non-linear words are pure]
-  \label{lem:non-linear-pure}
-  If $\npair{(w_1,w_2)} \in \lrv(H_\sigma,W)$ and ($\nonLinear{w_1}$ or $\nonLinear{w_2}$), then
-  $\npair{(w_1,w_2)} \in \lrv(H_\sigma,\purePart{W})$.
-\end{lemma}
-\begin{proof}
-  Follows easily by inspecting the definitions of $\lrv$, $\readCond[]{}$, $\addressable{}$, $\writeCond[]{}$, $\execCond[]{}$ and $\xReadCond[]{}$ and using Lemma~\ref{lem:purePart-idempotent}.
-\end{proof}
-
 \begin{lemma}[load from regular capability works]
@@
 \begin{enumproof}
-  \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T \step\failed$
+  \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T
+    \step[i] \failed$ for some $i$.
   \item \label{case:ftlr:halted} $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
@@
       \ints$.
-    \item TODO: other useful info
+    \item $\Phi_S \step[\gc] \Phi_S' \neq \failed$
+    \item $\Phi_T \step \Phi_T' \neq \failed$
+    \item $\Phi_T' = \updPc{\Phi_T\updReg{\rtmp{1}}{z}}$
+    \item $z \in \ints$
+    \item $\Phi_S(r_1) = \sealed{\sigma,c_1}$
+    \item $\Phi_S(r_2) = \sealed{\sigma,c_2}$
+    \item $\nonExec{c_2}$
+    \item $\Phi_S(\stk) = \stkptr{\rw, \baddr_\stk, \eaddr_\stk, \aaddr_\stk}$
+    \item $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$
+    \item $\Phi_S(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+    \item $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$
+    \item
+      \[
+        \begin{split}
+          \Phi_S'.\reg = \Phi_S.\reg
+          &[\pcreg \mapsto c_1] \\
+          &[\rdata \mapsto c_2] \\
+          &[\rstk \mapsto \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}] \\
+          &[\rretc \mapsto \sealed{\sigma',\retptrc{\baddr,\eaddr,\aaddr+\calllen}}]\\
+          &[\rretd \mapsto \sealed{\sigma',\retptrd{\aaddr_\stk,\eaddr_\stk}}] \\
+          &[r_1,r_2 \mapsto w_1,w_2] \\
+          &[\rtmp{1} \mapsto 0]
+        \end{split}
+      \]
+    \item $\Phi_S'.\mem = \Phi_S.\mem$
+    \item $\Phi_S'.\stk = ((\aaddr+\calllen),\ms_\stk |_{\aaddr_\stk,\eaddr_\stk}\update{\aaddr_\stk}{42}) :: \Phi_S.\stk$
+    \item $\Phi_S'.\ms_\stk = \Phi_S.\ms_\stk = \ms_\stk - \ms_\stk|_{\aaddr_\stk,\eaddr_\stk}$
+    \item $\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}$
+    \item $\sigma' = \sigma_\aaddr + \offsigma$
+    \item $\sigma_\aaddr \leq \sigma' \leq \sigma_\eaddr$
     \end{itemize}

### 2018-05-29 fully-abstract-comp/technical-report.tex
Fix splittability of readCond and writeCond and xReadCond.
Proposal for making execCond splittable.
Add some more lemmas about splittability, shrinkability and purity of permission-based conditions and the value relation.

@@
 %  \item Opaque capabilities do not hide the length of a program (especially an issue if we ever hope to have error recovery).
-  \item This is needed for correctness. See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
+  \item This is needed for correctness. %See subsection~\ref{subsec:capability-opacity} for an example it helps prevent.
   \item It is only used for the call instruction.
@@
         %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
-        \perm = \rx &\Rightarrow& \array[t]{l}\npair{(\{\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge\\
+        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{\lin,W} \wedge\\
                                   \npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}\endarray
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
       \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
@@
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
+      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
       \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
@@
 Note: this new version of execCond uses the expression relation for regular jumps since it expresses the validity of jumping to an address pointed to by an executable capability. 
-\[
-  \execCond{\lin,W} = \left\{ \npair{(\baddr,\eaddr)} \middle|
-    \begin{array}{l}
-      \forall W_l \ldotp\\
-      \quad ((W_l = \purePart{W} \wedge \lin = \normal) \vee (W_l = W \wedge \lin = \linear)) \wedge\\
-      \quad \forall n' < n, W' \future W_l, \aaddr \in [\baddr',\eaddr'] \subseteq [\baddr,\eaddr] \ldotp\\
-      \qquad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+\begin{align*}
+  \execCond{\lin,W} &=
+  \left\{ \npair{A} \middle|
+    \begin{array}{l}
+      \forall n' < n, W' \future \purePart{W}, \aaddr \in [\baddr',\eaddr'] \subseteq A \ldotp\\
+      \quad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+    \end{array}
+    \right\}\\
+  &\cup\left\{ \npair{A} \middle|
+    \begin{array}{l}
+      \lin = \linear \wedge\\
+      \exists W_a : A \fun \World \wedge W = \oplus_{a\in[\baddr,\eaddr]} W_a(a)\wedge\\
+      \forall n' < n, \aaddr\in[\baddr',\eaddr'] \subseteq A, W' \future \oplus_{a\in[\baddr',\eaddr']} W_a(a), \ldotp\\
+      \quad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
     \end{array}
     \right\}
-\]
+\end{align*}
 
@@
 
-\begin{lemma}[readCond shrinkable]
-  \label{lem:readCond-shrinkable}
-  If $\npair{A} \in \readCond{\lin,W}$  and $A' \subseteq A$, then
-  $\npair{A'} \in \readCond{\lin,W}$.
+\begin{lemma}[non-linear words are pure]
+  \label{lem:non-linear-pure}
+  \begin{itemize}
+  \item If $\npair{A} \in \readCond{\normal,W}$, then $\npair{A} \in \readCond{}(\normal, \purePart{W} )$.
+  \item If $\npair{A} \in \writeCond{\normal,W}$, then $\npair{A} \in \writeCond{}(\normal,\purePart{W})$
+  \item If $\npair{A} \in \execCond{\normal,W}$, then $\npair{A} \in \execCond{}(\normal,\purePart{W})$
+  \item If $\npair{(w_1,w_2)} \in \lrv(H_\sigma,W)$ and ($\nonLinear{w_1}$ or
+    $\nonLinear{w_2}$), then
+    $\npair{(w_1,w_2)} \in \lrv(H_\sigma,\purePart{W})$.
+  \end{itemize}
 \end{lemma}
 \begin{proof}
-  Follows easily from the definition.
+  Follows easily by inspecting the definitions of $\lrv$, $\readCond[]{}$, $\addressable{}$, $\writeCond[]{}$, $\execCond[]{}$ and $\xReadCond[]{}$ and using Lemma~\ref{lem:purePart-idempotent}.
 \end{proof}
 
-\begin{lemma}[readXCond shrinkable]
-  \label{lem:readXCond-shrinkable}
-  If $\npair{A} \in \xReadCond{\lin,W}$  and $A' \subseteq A$, then
-  $\npair{A'} \in \xReadCond{\lin,W}$.
+\begin{lemma}[permission-based conditions splittable]
+  \label{lem:conds-splittable}
+  If $A = [\baddr_1,\eaddr_1] \uplus [\baddr_2,\eaddr_2]$, then
+  \begin{itemize}
+  \item If $\npair{A} \in \readCond{\lin,W}$, then there exists $W_1, W_2$ such that $W = W_1\oplus W_2$ such that $\npair{[\baddr_1,\eaddr_1]} \in \readCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \readCond{\lin,W_2}$.
+  \item If $\npair{A} \in \writeCond{\lin,W}$, then there exists $W_1, W_2$ such that $W = W_1\oplus W_2$ such that $\npair{[\baddr_1,\eaddr_1]} \in \writeCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \writeCond{\lin,W_2}$.
+  \item If $\npair{A} \in \execCond{\lin,W}$, then there exists $W_1, W_2$ such that $W = W_1\oplus W_2$ such that $\npair{[\baddr_1,\eaddr_1]} \in \execCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \execCond{\lin,W_2}$.
+  \end{itemize}
 \end{lemma}
 \begin{proof}
-  Follows easily from the definition.
+  TODO
+\end{proof}
+
+\begin{lemma}[permission-based conditions shrinkable]
+  \label{lem:conds-shrinkable}
+  \begin{itemize}
+  \item If $\npair{A} \in \readCond{\lin,W}$  and $A' \subseteq A$, then
+    $\npair{A'} \in \readCond{\lin,W}$.
+  \item If $\npair{A} \in \writeCond{\lin,W}$  and $A' \subseteq A$, then
+    $\npair{A'} \in \readCond{\lin,W}$.
+  \item If $\npair{A} \in \xReadCond{\lin,W}$  and $A' \subseteq A$, then
+    $\npair{A'} \in \xReadCond{\lin,W}$.
+  \item If $\npair{A} \in \execCond{\lin,W}$  and $A' \subseteq A$, then
+    $\npair{A'} \in \execCond{\lin,W}$.
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Follows easily from the definition. TODO: check after update
 \end{proof}
@@
       Then we need to show that $\npair[n']{(w',w')} \in \lre(W')$.
-      This now follows immediately from the (regular) FTLR (Theorem~\ref{thm:ftlr}), using the two above points, Lemma~\ref{lem:downwards-closed} and Lemmas~\ref{lem:readCond-shrinkable} and~\ref{lem:readXCond-shrinkable}.
+      This now follows immediately from the (regular) FTLR (Theorem~\ref{thm:ftlr}), using the two above points, Lemma~\ref{lem:downwards-closed} and Lemma~\ref{lem:conds-shrinkable}.
     \end{itemize}

### 2018-05-29 fully-abstract-comp/technical-report.tex
Typo.
Proofs for purity, shrinkability and splittability lemmas.

@@
       \exists W_a : A \fun \World \wedge W = \oplus_{a\in[\baddr,\eaddr]} W_a(a)\wedge\\
-      \forall n' < n, \aaddr\in[\baddr',\eaddr'] \subseteq A, W' \future \oplus_{a\in[\baddr',\eaddr']} W_a(a), \ldotp\\
+      \forall n' < n, \aaddr\in[\baddr',\eaddr'] \subseteq A, W' \future \oplus_{a\in[\baddr',\eaddr']} W_a(a) \ldotp\\
       \quad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
@@
   \label{lem:non-linear-pure}
+
   \begin{itemize}
@@
 
+\begin{lemma}[permission-based conditions shrinkable]
+  \label{lem:conds-shrinkable}
+  \begin{itemize}
+  \item If $\npair{A} \in \readCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
+    $\npair{A'} \in \readCond{\lin,W}$.
+  \item If $\npair{A} \in \writeCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
+    $\npair{A'} \in \writeCond{\lin,W}$.
+  \item If $\npair{A} \in \xReadCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
+    $\npair{A'} \in \xReadCond{\lin,W}$.
+  \item If $\npair{A} \in \execCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
+    $\npair{A'} \in \execCond{\lin,W}$.
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+  In the case for $\execCond$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
+  It suffices construct new $W_a'$ by taking $W_a$ but adding $\oplus_{a\in(A\setminus A')}$ to an $W_a(a)$ for an arbitrary $a \in A'$ to make this work.
+\end{proof}
+
 \begin{lemma}[permission-based conditions splittable]
@@
 \begin{proof}
-  TODO
-\end{proof}
-
-\begin{lemma}[permission-based conditions shrinkable]
-  \label{lem:conds-shrinkable}
-  \begin{itemize}
-  \item If $\npair{A} \in \readCond{\lin,W}$  and $A' \subseteq A$, then
-    $\npair{A'} \in \readCond{\lin,W}$.
-  \item If $\npair{A} \in \writeCond{\lin,W}$  and $A' \subseteq A$, then
-    $\npair{A'} \in \readCond{\lin,W}$.
-  \item If $\npair{A} \in \xReadCond{\lin,W}$  and $A' \subseteq A$, then
-    $\npair{A'} \in \xReadCond{\lin,W}$.
-  \item If $\npair{A} \in \execCond{\lin,W}$  and $A' \subseteq A$, then
-    $\npair{A'} \in \execCond{\lin,W}$.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  Follows easily from the definition. TODO: check after update
+  If $\lin = \normal$, then Lemma~\ref{lem:non-linear-pure} tells us that it suffices to consider pure worlds $W = \purePart{W}$ and Lemma~\ref{lem:purePart-duplicable} tells us that then $W = W \oplus W$.
+  The results then follow from the previous Lemma~\ref{lem:conds-shrinkable}.
+
+  As such, we can limit ourselves to the case where $\lin = \linear$.
+
+  The read- and write-conditions then require separate islands for each individual address in $A$ and we can distribute ownership of those islands according to whether those addresses are in $[\baddr_1,\eaddr_1]$ or $[\baddr_2,\eaddr_2]$, i.e.
+make the island $\pwheap(r)$ with $R(r) = \{a\}$ $\spatialo$ in $W_1$ and $\spatial$ in $W_2$ if $a \in [\baddr_1,\eaddr_1]$ and vice versa.
+  It is then easy to check that our results hold.
+
+  For the execute condition, we get a partition of the world into $W = \oplus_{a \in A} W_a(a)$ and we can define $W_1 \isdef \oplus_{a\in[\baddr_1,\eaddr_1]} W_a(a)$ and likewise for $W_2$.
+  The results then follow easily from the definitions.
 \end{proof}

### 2018-05-30 fully-abstract-comp/technical-report.tex
Updated xjumpResult to not give undefined and changed scall semantics to use xjumpResult.
Fix latex compilation.

@@
     \\
-    \tundefined & \totherwise
+    \failed & \totherwise
   \end{cases}
@@
 We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}}
-\begin{align*}
-  \sem{\scall{\offpc,\offsigma}{\src{r_1}}{\src{r_2}}}(\Phi) = & 
-                                               \begin{cases}
-                                                 \Phi'
-                                                 \arraycolsep=0pt
-                                                 \begin{array}[t]{rl}
-                                                   &\updReg{r_1}{w_1} \\
+
+\begin{multline*}
+  \sem{\scall{\offpc,\offsigma}{\src{r_1}}{\src{r_2}}}(\Phi) = \\
+  \begin{cases}
+    \mathit{xjumpResult}\left(c_1,c_2, 
+     \arraycolsep=0pt
+    \array[]{rl}
+      \Phi'&\updReg{r_1}{w_1} \\
       &\updReg{r_2}{w_2} \\
-                                                        &\updReg{\pcreg}{c_1}\\
-                                                        &\updReg{\rdata}{c_2}\\
       &\updReg{\rstk}{c_\stk}\\
@@
       &\updReg{\rtmp{1}}{0}
-                                                 \end{array}
-                                                 & 
-                                                 \arraycolsep=0pt
+      \endarray
+     \right)
+    & \arraycolsep=0pt
     \begin{array}[t]{l}
@@
   \end{cases}
-\end{align*}
+\end{multline*}
 \lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$.
@@
 
-  For the execute condition, we get a partition of the world into $W = \oplus_{a \in A} W_a(a)$ and we can define $W_1 \isdef \oplus_{a\in[\baddr_1,\eaddr_1]} W_a(a)$ and likewise for $W_2$.
+  For the execute condition, we get a partition of the world into $W = \oplus_{a \in A} W_a(a)$ and we can define $W_1 = \oplus_{a\in[\baddr_1,\eaddr_1]} W_a(a)$ and likewise for $W_2$.
   The results then follow easily from the definitions.

### 2018-05-30 fully-abstract-comp/technical-report.tex
Finish a proof about safety of untrusted code pointers (almost).

@@
       \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
-      \exists S : A \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
+      \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
       \quad \forall \aaddr \in A, n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
@@
 
+\begin{lemma}[Untrusted code regions' seals specify value safety]
+  \label{lem:untrusted-codereg-sealed-vals-safe}
+  If
+  \begin{itemize}
+  \item $\dom(\code) \mathrel{\#} \ta$
+  \item $\sigma \in \sigcloss$
+  \end{itemize}
+  Then
+  \begin{equation*}
+    H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \nequal \lrv
+  \end{equation*}
+\end{lemma}
+\begin{proof}
+  Follows easily by definition.
+\end{proof}
+
 \begin{lemma}[Code region for untrusted components stronger than standard safe memory region]
@@
   \begin{itemize}
-  \item $\dom(\mscode) \mathrel{\#} \ta$
-  \end{itemize}
-  then $\codereg{\sigrets,\sigcloss,\mscode,\gc} \nsubeq \stdreg{R(r),\gc}{\pur}$ 
+  \item $\emptyset\neq\dom(\mscode) \mathrel{\#} \ta$
+  \end{itemize}
+  then $\codereg{\sigrets,\sigcloss,\mscode,\gc} \nsubeq \stdreg{\dom(\mscode),\gc}{\pur}$ 
 \end{lemma}
 \begin{proof}
-  TODO
+  Take $\hat{W}$, then we need to prove that
+  \begin{equation*}
+    H^\mathrm{code}\ ;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubeq H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
+  \end{equation*}
+  So take $\npair{\ms_S,\ms_T} \in H^\mathrm{code}\ ;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W}$, then we know that
+  \begin{itemize}
+  \item $\ms_S = \ms_T = \mscode$ 
+  \item $\dom(\mscode) =[\aaddr,\baddr]$
+  \item $\dom(\mscode) \subseteq \ta \vee (\dom(\mscode)\mathrel{\#} \ta \wedge \sigrets = \emptyset)$, but we know that $\emptyset \neq\dom(\mscode) \mathrel{\#} \ta$, so we get that $\sigrets = \emptyset$.
+  \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$
+  \end{itemize}
+
+  By definition of $H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}$, we need to prove that
+  \begin{itemize}
+  \item $\dom(\mscode) = \dom(\mscode) = \dom(\mscode)$
+  \item there exists a $S : \dom(\mscode) \fun \World$ with $\xi(\hat{W}) = \oplus_{\aaddr\in\dom(\mscode)} S(\aaddr)$ and for all $\aaddr \in \dom(\mscode)$, $n'<n$, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$
+  \end{itemize}
+
+  We take $S$ to map every address to $\purePart{\hat{W}}$, except for a single one that we map to $\hat{W}$.
+  Take $\aaddr\in\dom(\mscode)$, $n'<n$.
+  By Lemmas~\ref{lem:purePart-duplicable}, \ref{lem:oplus-future} and \ref{lem:monotonicity}, it now suffices to prove that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(\purePart{\hat{W}})$.
+
+  From $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$ and the fact that $\sigrets = \emptyset$, we know that for any $\aaddr\in\dom(\mscode)$, we have that $\emptyset,\emptyset,\sigcloss \vdash_{\mathrm{comp-code}} \mscode,\aaddr$.
+  By definition, this last judgement implies that one of the following cases holds:
+  \begin{itemize}
+  \item $\mscode(a) \in\ints$ and an omitted extra fact.  It follows trivially that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(\purePart{\hat{W}})$.
+  \item 
+    $\mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ and $[\sigma_\baddr,\sigma_\eaddr] \subseteq \sigcloss$.
+
+    By definition of $\lrv$, it suffices to show that $\exists r \in
+    \dom(\pwheap)$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \;
+    \sigma' \nequal \lrv$.
+
+    \dominique{TODO: we need to know that our code region is in $\hat{W}$?}
+
+    Then we can use Lemma~\ref{lem:untrusted-codereg-sealed-vals-safe}.
+  \end{itemize}
 \end{proof}

### 2018-05-30 fully-abstract-comp/technical-report.tex
Finish proof by adding an extra technical requirement to codereg.

@@
     \dom(\code) \subseteq \ta \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset) \wedge \\
-    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code 
+      \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
+      \forall a \in \dom(\code), n' < n \ldotp\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})
     \end{array}
@@
   \item $\dom(\mscode) \subseteq \ta \vee (\dom(\mscode)\mathrel{\#} \ta \wedge \sigrets = \emptyset)$, but we know that $\emptyset \neq\dom(\mscode) \mathrel{\#} \ta$, so we get that $\sigrets = \emptyset$.
-  \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$
+  \item For all $a \in \dom(\code)$, $n'<n$, $\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$
   \end{itemize}
@@
   We take $S$ to map every address to $\purePart{\hat{W}}$, except for a single one that we map to $\hat{W}$.
-  Take $\aaddr\in\dom(\mscode)$, $n'<n$.
-  By Lemmas~\ref{lem:purePart-duplicable}, \ref{lem:oplus-future} and \ref{lem:monotonicity}, it now suffices to prove that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(\purePart{\hat{W}})$.
-
-  From $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$ and the fact that $\sigrets = \emptyset$, we know that for any $\aaddr\in\dom(\mscode)$, we have that $\emptyset,\emptyset,\sigcloss \vdash_{\mathrm{comp-code}} \mscode,\aaddr$.
-  By definition, this last judgement implies that one of the following cases holds:
-  \begin{itemize}
-  \item $\mscode(a) \in\ints$ and an omitted extra fact.  It follows trivially that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(\purePart{\hat{W}})$.
-  \item 
-    $\mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ and $[\sigma_\baddr,\sigma_\eaddr] \subseteq \sigcloss$.
-
-    By definition of $\lrv$, it suffices to show that $\exists r \in
-    \dom(\pwheap)$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \;
-    \sigma' \nequal \lrv$.
-
-    \dominique{TODO: we need to know that our code region is in $\hat{W}$?}
-
-    Then we can use Lemma~\ref{lem:untrusted-codereg-sealed-vals-safe}.
-  \end{itemize}
+  The result then follows by Lemmas~\ref{lem:purePart-duplicable}, \ref{lem:oplus-future} and \ref{lem:monotonicity} and the above fact that for all $a \in \dom(\code)$, $n'<n$, $\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$.
 \end{proof}

### 2018-05-30 fully-abstract-comp/technical-report.tex
Finish a minor TODO using the shrinkability lemma.

@@
     \begin{itemize}
-    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
+    \item $\npair[n']{(\vsc,\vsc)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
 
@@
    \item 
-     For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+     For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$, we have that
      \begin{equation*}
@@
      \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
-   \end{itemize}
      TODO
+   \end{itemize}
+
   \item $w = ((\perm,\lin),\baddr,\eaddr,\aaddr)$,
@@
     $\lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}$:
-    TODO
+
+    By definition of $\lrv$, it suffices to prove that $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ and $\npair{[\baddr,\eaddr]}\in\writeCond{\lin,W}$.
+    The result then follows easily from the assumptions, with Lemma~\ref{lem:conds-shrinkable}.
 

### 2018-05-31 fully-abstract-comp/technical-report.tex
Fixed permission based stack conditions.
Typos and formatting.

@@
       c_2 \neq \retptrd(\_) \tand \\
-      \nonExec{c_\data}
+      \nonExec{c_2}
     \end{array}
@@
 \end{align*}
-
+\clearpage
 We define inductively when a component is valid ($\vdash \var{comp}$) by the below inference rules:
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
-      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
     \begin{array}{l}
-      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
       \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \forall r \in S \ldotp |R(r)| = 1\\
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
-      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
+      \qquad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
-      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
 \[
-  \stackWriteCond{W} = \left\{ \npair{[\baddr,\eaddr]}) \middle| 
-    \begin{array}{l}
-      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
+  \stackWriteCond{W} = \left\{ \npair{A}) \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
       \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr] \wedge \\
-      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
+      \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}

### 2018-05-31 fully-abstract-comp/ftlr-proof-cases.tex
Updated notation for n sub and super sets as they are not really nsub/sup

@@
     \pwheap}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwheap(r)
-  \nsubeq \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
+  \nsubsim \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
 
@@
   \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
-  Because $\pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
-  also $\pwheap[W_\var{heap}](r) \nsubeq \stdreg{R(r),\gc}{\pur}$. This
+  Because $\pwheap(r) \nsubsim \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
+  also $\pwheap[W_\var{heap}](r) \nsubsim \stdreg{R(r),\gc}{\pur}$. This
   means that also $\npair{R_\ms(r)} \in
-  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined $\nsubeq$ somewhere?}
+  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined
+    $\nsubsim$ somewhere?} \lau{Yes, it is in subsection~\ref{subsec:worlds}}
 
@@
   we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
-  for all $r \in S'$, $\pwheap(r) \nsupeq \stdreg{R'(r),\gc}{\pur}$ and
+  for all $r \in S'$, $\pwheap(r) \nsupsim \stdreg{R'(r),\gc}{\pur}$ and
   $\pwheap(r) \text{ is address-stratified}$.
@@
   Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwheap[W_\var{heap}](r') \nsupeq \stdreg{R'(r'),\gc}{\pur}$ and
+  $\pwheap[W_\var{heap}](r') \nsupsim \stdreg{R'(r'),\gc}{\pur}$ and
   $\pwheap[W_\var{heap}](r') \text{ is address-stratified}$.
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupeq \stdreg{R'(r),\gc}{\pur}$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupsim \stdreg{R'(r),\gc}{\pur}$.
 
@@
     \pwfree}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwfree(r)
-  \nsubeq \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
+  \nsubsim \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
 
@@
 
-  We have that $r \in \addressable{\lin, \pwfree} \subseteq
-  \activeReg{\pwfree[W_\var{free\_stack}]}$, so $\npair{R_\ms(r)} \in
-  \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
-  Because $\pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{free\_stack} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
-  also $\pwfree[W_\var{free\_stack}](r) \nsubeq \stdreg{R(r),\gc}{\pur}$. This
-  means that also $\npair{R_\ms(r)} \in
-  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined $\nsubeq$ somewhere?}
+  We have that $r \in \addressable{\lin, \pwfree} \subseteq \activeReg{\pwfree[W_\var{free\_stack}]}$, so $\npair{R_\ms(r)} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
+  Because $\pwfree(r) \nsubsim \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{free\_stack} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that also $\pwfree[W_\var{free\_stack}](r) \nsubsim \stdreg{R(r),\gc}{\pur}$.
+  This means that also $\npair{R_\ms(r)} \in H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined $\nsubsim$ somewhere?}
+  \lau{Yes, it is in subsection~\ref{subsec:worlds}}
 
@@
   we get an $S'\subseteq \addressable{\lin, \pwfree} \subseteq \activeReg{\pwheap[W_\var{free\_stack}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ and
-  for all $r \in S'$, $\pwfree(r) \nsupeq \stdreg{R'(r),\gc}{\pur}$ and
+  for all $r \in S'$, $\pwfree(r) \nsupsim \stdreg{R'(r),\gc}{\pur}$ and
   $\pwfree(r) \text{ is address-stratified}$.
@@
   Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwfree[W_\var{free\_stack}](r') \nsupeq \stdreg{R'(r'),\gc}{\pur}$ and
+  $\pwfree[W_\var{free\_stack}](r') \nsupsim \stdreg{R'(r'),\gc}{\pur}$ and
   $\pwfree[W_\var{free\_stack}](r') \text{ is address-stratified}$.
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r) \nsupeq \stdreg{R'(r),\gc}{\pur}$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r) \nsupsim \stdreg{R'(r),\gc}{\pur}$.
 

### 2018-05-31 fully-abstract-comp/technical-report.tex
Updated notation for n sub and super sets as they are not really nsub/sup

@@
 \subsection{Worlds}
+\label{subsec:worlds}
 \[
@@
 
-\dominique{Remaining question: what exactly does $\nsubeq$ on a UPred mean?  Does it mean subset ``up to n (including)'' or ``up to n-1 (including)''?
+\dominique{Remaining question: what exactly does $\nsubsim$ on a UPred mean?  Does it mean subset ``up to n (including)'' or ``up to n-1 (including)''?
   Check what we can require without breaking contractiveness of recursive definition of value relation and/or what we did in previous paper and/or what we need in FTLR proof.
@@
 \begin{mathpar}
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
-            { (v,H) \nsubeq (v',H') }
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+            { (v,H) \nsubsim (v',H') }
 \and
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
-            { (\pure,H,H_\sigma) \nsubeq (v',H') }
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+            { (\pure,H,H_\sigma) \nsubsim (v',H') }
 \and
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
-            { (v,H) \nsubeq (\pure,H',H_\sigma) }
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+            { (v,H) \nsubsim (\pure,H',H_\sigma) }
 \and
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubeq H' \; \hat{W}}
-            { (\pure,H,H_\sigma) \nsubeq (\pure,H',H_\sigma') }
+  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+            { (\pure,H,H_\sigma) \nsubsim (\pure,H',H_\sigma') }
 \end{mathpar}
-We define $n$-equality of regions $\iota \nequal \iota'$ as the symmetric closure, i.e. $\iota \nequal \iota'$ iff $\iota \nsubeq \iota'$ and $\iota \nsupeq \iota'$.
+We define $n$-equality of regions $\iota \nsim \iota'$ as the symmetric closure, i.e. $\iota \nequal \iota'$ iff $\iota \nsubsim \iota'$ and $\iota \nsupsim \iota'$.
 
@@
       \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad \forall r \in S \ldotp \pwheap(r) \nsubsim \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsubsim \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
+      \qquad \forall r \in S \ldotp  \pwheap(r) \nsim \codereg{\_,\_,M(r),\gc}
     \end{array}
@@
       \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r) \nsupsim \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
       \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsupsim \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}
@@
 \]
-\lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubeq$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
+\lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubsim$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
 where $H^{\mathrm{std},\square}_A$ is defined as follows:
@@
         \dom(\code) \subseteq \ta \tand\\
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nsim (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
@@
   \begin{equation*}
-    H^\mathrm{code}\ ;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubeq H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
+    H^\mathrm{code}\ ;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubsim H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
   \end{equation*}
@@
 \section{Lemmas}
+
+\begin{lemma}
+  \label{lem:nsub-and-nsup-std}
+  If $\iota \nsubsim \stdreg{A,\gc}{v}$ and $\iota \nsupsim \stdreg{A,\gc}{v}$, then $\iota \nsim {A,\gc}{v}$.
+\end{lemma}
+\begin{proof}
+  By definition.
+\end{proof}
 

### 2018-05-31 fully-abstract-comp/technical-report.tex
Small correction in FTLR statement and a missed nsubeq

@@
     \item $\trust = \trusted$
-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
+    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}\updReg{\pcreg}{((\rx,\lin),\baddr,\eaddr,\aaddr)}, \src{\stk}, \src{\ms_\stk})$ and
       \begin{itemize}           
@@
   \end{itemize}
-  then $\codereg{\sigrets,\sigcloss,\mscode,\gc} \nsubeq \stdreg{\dom(\mscode),\gc}{\pur}$ 
+  then $\codereg{\sigrets,\sigcloss,\mscode,\gc} \nsubsim \stdreg{\dom(\mscode),\gc}{\pur}$ 
 \end{lemma}

### 2018-06-04 fully-abstract-comp/technical-report.tex
Fix "world specifies component" relation and prove FTLR for components (modulo 2 TODOs).

--- 2018-05-31 fully-abstract-comp/technical-report.tex
@@
       c_2 \neq \retptrd(\_) \tand \\
-      \nonExec{c_2}
+      \nonExec{c_\data}
     \end{array}
@@
 \begin{align*}
-  \var{comp}_0 &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)\\
+  \var{comp}_0 &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)\\
   \var{comp} &\mathrel{::=} \var{comp}_0\\
@@
 \end{align*}
-\clearpage
+
 We define inductively when a component is valid ($\vdash \var{comp}$) by the below inference rules:
@@
   }
-  \and
+\end{mathpar}
+
+\begin{mathpar}
   \inference{
@@
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
-    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
+    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
+    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
     \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
@@
   }{
-    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
   }
@@
   \inference{
-    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)\\
+    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)\\
     \vdash \var{comp}_0 & c_{\mathrm{main},c}, c_{\mathrm{main},d} \in \overline{c_{\mathrm{export}}}
@@
   \inference{
-    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
-    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
-    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
+    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
+    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
+    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
     \mscode[3] = \mscode[1] \uplus \mscode[2] &
@@
     \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
-    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
-  }{
+    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
+    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}
+  } {
     \var{comp}_3 = \var{comp}_1 \bowtie \var{comp}_2
@@
     }{
-      ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
+      ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
     }
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
+      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
-      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
+      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
+      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \quad \forall r \in S \ldotp |R(r)| = 1 \\
+      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}
@@
   \begin{equation*}
-    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
   \end{equation*}
   (written as $W \vdash_\gc \var{comp}_0$)
-  iff
+  iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that for all $i$,
+  \begin{itemize}
+  \item 
     \begin{equation*}
-    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+      \vdash (\code_i,\msdata[i],\overline{c_{\mathrm{import},i}},\overline{c_{\mathrm{export},i}},\sigrets_i,\sigcloss_i,A_{\linear,i})
     \end{equation*}
-  and there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that
-  \begin{itemize}
+  \item $\mscode = \biguplus_i \code_i$
+  \item $\msdata = \biguplus_i \msdata[i]$
+  \item $\overline{c_{\mathrm{export}}} = \biguplus_i \overline{c_{\mathrm{export},i}}$
+  \item $\overline{c_{\mathrm{import}}} = \biguplus_i \overline{c_{\mathrm{import},i}} \setminus \overline{c_{\mathrm{export}}}$
+  \item $\sigrets = \biguplus \sigrets_i$
+  \item $\sigcloss = \biguplus \sigcloss_i$
+  \item $A_\linear = \biguplus A_{\linear,i}$
   \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
-  \item $\mscode = \biguplus_i \code_i$
   \item For all $i$, one of the following holds:
@@
     \end{itemize}
-  \item $\sigrets = \biguplus_i \sigrets_i$
-  \item $\sigcloss = \biguplus_i \sigcloss_i$.
-  \end{itemize}
-  and there exist $r_1 \cdots r_m \in \dom(\pwheap)$ such that
-  \begin{itemize}
-  \item $\pwheap(r_i) = \stdreg{A_i,\gc}{\lin}$ with $\lin = \pure$ or $\lin = \spatialo$
-  \item $\dom(\msdata) = \biguplus_i A_i$
+  \end{itemize}
+  and there exist $r_{\mathrm{addr}} : \dom(\msdata) \mapsto \dom(\pwheap)$ such that
+  \begin{itemize}
+  \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) or ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
   \end{itemize}
@@
   \item \begin{itemize}
-    \item $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$
+    \item $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss,\_)$
     \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
@@
       \end{itemize}
+      \dominique{something missing here: pc?}
     \end{itemize}
@@
   By definition of $\bowtie$, we have that the following hold:
-  \begin{align*}
-    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
-    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
-    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
-    \mscode[3] = \mscode[1] \uplus \mscode[2] &
-    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
-    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
-    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
-    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
-    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
-  \end{align*}
+  \begin{itemize}
+  \item $\var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1],A_{\linear,1})$
+  \item $\var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2],A_{\linear,2})$
+  \item $\var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3],A_{\linear,3})$
+  \item $\mscode[3] = \mscode[1] \uplus \mscode[2]$
+  \item $\msdata[3] = \msdata[1] \uplus \msdata[2]$
+  \item $\overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}$
+  \item $\overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}$
+  \item $\sigrets[3] = \sigrets[1] \uplus \sigrets[2]$
+  \item $\sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]$
+  \item $A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}$
+  \end{itemize}
 
@@
   \begin{itemize}
-  \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item $\npair{\dom(\mscode)}\in\xReadCond{\normal,W}$
   \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
@@
     \item $\npair{[b,e]}\in\xReadCond{\normal,W}$:
-      follows from the definition of $\xReadCond$, together with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+      by assumption and Lemma~\ref{lem:conds-shrinkable}.
     \item $\npair{[b,e]}\in\execCond{\normal,W}$:
@@
   \item $\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$:
-    By definition of $\lrvg{\trusted}(W)$, it suffices to show that
-    $\npair{[b,e]}\in\xReadCond{\normal,W}$.
-
-    This follows immediately from the definition of $\xReadCond$, together with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+    By definition of $\lrvg{\trusted}(W)$ and Lemma~\ref{lem:conds-shrinkable} with the assumption that $\npair{\dom(\mscode)}\in\xReadCond{\normal,W}$.
   \end{itemize}
@@
 \begin{lemma}[FTLR for component data-values]
+  \label{lem:ftlr-comp-value}
   If
@@
   \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
-  \item $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$
-  \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(W)$ for all $n' < n$
+  \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$
+  \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(\purePart{W})$ for all $n' < n$
   \item $w \not\in \overline{c_{\mathrm{import}}}$
@@
 
-     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs} that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(W)$, using Lemma~\ref{lem:downwards-closed}.
+     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs} that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(W)$, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
 
@@
   \begin{equation*}
-    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss)
+    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
   \end{equation*}
 
-  We know that $W \vdash \var{comp}$ by assumption, so it remains to prove that 
-  for all  $W' \future W$ with 
+  We know that $W \vdash \var{comp}$ by assumption, so take $W' \future W$ and
   \begin{equation*}
@@
   \end{equation*}
-  we have that
+  then it remains to prove that
   \begin{equation*}
@@
 
-  TODO
-  \dominique{Something missing: we should be able to subdivide the world according to how linear authority is distributed in the component well-formedness judgement, but the world-specifies-component relation does not give us this.}
+  We prove by general induction on $n$ that this is the case for all $\overline{c_{\mathrm{export}}}$, i.e. assume that for all $n' < n$,
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+  \end{equation*}
+
+  Because $W \vdash \var{comp}$, for every $c_{\mathrm{export}} \in \overline{c_{\mathrm{export}}}$, there must exist an $i$ such that
+  \begin{itemize}
+  \item
+    \begin{equation*}
+      \var{comp}_i = (\code_i,\msdata[i],\overline{c_{\mathrm{import},i}},\overline{c_{\mathrm{export},i}},\sigrets_i,\sigcloss_i,A_{\linear,i})
+    \end{equation*}
+  \item $\vdash \var{comp}_i$
+  \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
+  \item One of the following holds:
+    \begin{itemize}
+    \item $\dom(\code_i) \subseteq \ta$
+    \item $\dom(\code_i) \mathrel{\#} \ta$
+    \end{itemize}
+  \item $c_{\mathrm{export}} \in \overline{c_{\mathrm{export},i}}$
+  \end{itemize}
+
+  It follows easily that for any $n$,
+  \begin{itemize}
+  \item $\npair{A_{\linear,i}} \in \readCond{\linear,W}$,
+  \item $\npair{A_{\linear,i}} \in \writeCond{\linear,W}$,
+  \item $\npair{A_{\mathrm{non-linear},i}} \in \readCond{\normal,W}$ and
+  \item $\npair{A_{\mathrm{non-linear},i}} \in \writeCond{\normal,W}$
+  \end{itemize}
+  with $A_{\mathrm{non-linear},i} = \dom(\msdata[i]) \setminus A_{\linear,i}$
+
+  We have that the component is well-formed: $\vdash \var{comp}_i$, which implies that
+  \begin{equation*}
+    \dom(\mscode[i]), \emptyset, A_{\mathrm{non-linear},i}, \overline{c_{\mathrm{import},i}}, \sigrets_i,\sigcloss_i \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}
+  \end{equation*}
+  and $\overline{c_{\mathrm{export},i}} \mathrel{\#} \overline{c_{\mathrm{import},i}}$.
+
+  Using Lemma~\ref{lem:ftlr-comp-value} and the above results, it then remains to prove that
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import},i},c_{\mathrm{import},i})}} \in \lrv(\purePart{W}) \text{ for all } n' < n
+  \end{equation*}
+
+  From $W \vdash \var{comp}$, we know that $\overline{c_{\mathrm{import},i}} \subseteq \overline{c_{\mathrm{import}}} \cup \overline{c_{\mathrm{export}}}$.
+  For the former, we know by assumption that 
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
+  \end{equation*}
+  and for the latter, our induction hypothesis tells us that
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
+  \end{equation*}
 \end{proof}

### 2018-06-05 fully-abstract-comp/technical-report.tex
Discharge some simple proofs.
Start on proof of compatibility of initial configuration.

@@
     \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
-    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}
+    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
+    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]\\
   } {
@@
 \begin{proof}
-  TODO
+  Follows by induction on the component validity judgement and definition-chasing.
+  The most interesting point is related to imports that are resolved from the other component during linking, but we know that those exports satisfy the $\mathrm{comp-value}$ judgement.
 \end{proof}
@@
 \begin{proof}
-  TODO
+  Follows easily from the definitions.
 \end{proof}
@@
   \item $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$
-  \item $\var{comp}_S\rightsquigarrow \Phi_S$
-  \item $\var{comp}_T \rightsquigarrow \Phi_T$,
-  \end{itemize}
+  \item $\var{comp} \rightsquigarrow \Phi$,
+  \item TODO: no stack regions yet?
+  \end{itemize}
+  then $\exists W' \future W$ such that
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Take $\var{comp} = ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
+  Then we know from $\var{comp} \rightsquigarrow \Phi$ that
+  \begin{itemize}
+  \item $\reg(\pcreg) = c_{\mathrm{main},c}$
+  \item $\reg(\rdata) = c_{\mathrm{main},d}$ 
+  \item $\src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}}$ 
+  \item $\trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)}$ 
+  \item $\reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0$
+  \item $\range{\ms_\stk} = \{0\}$
+  \item $\mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk}$
+  \item $[\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata))$
+  \item $\overline{c_{\mathrm{import}}} = \emptyset$
+  \end{itemize}
+
+  From $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$, we know that
+  $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and 
+  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
+  From this, it follows that $W \vdash \var{comp}$ and for all  $W' \future W$, if
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
+  \end{equation*}
   then
-  \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W)$
-  \end{itemize}
-\end{lemma}
+  \begin{equation*}
+    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+  \end{equation*}
+  Since $\overline{c_{\mathrm{import}}} = \emptyset$, the former holds vacuously and the latter follows for every $W' \future W$.
+
+  Now take $\overline{r_a} \not\in\dom(\pwfree)$ and 
+  \begin{align*}
+    W' &= W\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\pur}}}_{a \in [\baddr_\stk,\eaddr_\stk]}\\
+   W_\pcreg &= \purePart{W'}\\
+    W_R &= \purePart{W}\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\spatialo}}}_{a \in [\baddr_\stk,\eaddr_\stk]}\\
+    W_M &= W\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\spatial}}}_{a \in [\baddr_\stk,\eaddr_\stk]}
+  \end{align*}
+  Then $W' = W_M \oplus W_R \oplus W_\pcreg$ and by definition of $\lrec$, it suffices to prove that 
+  \begin{itemize}
+  \item $\reg_S(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr)$:
+    TODO
+  \item $(\trust = \trusted \text{ iff } \aaddr \in \ta)$:
+    TODO
+  \item $\npair{( \reg_S(\pcreg),\reg_T(\pcreg) )} \in \lrvg{\trust}(W_\pcreg)$:
+    TODO
+  \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
+    TODO
+  \item $\npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$:
+    TODO
+  \end{itemize}
+
+\end{proof}
 

### 2018-06-05 fully-abstract-comp/technical-report.tex
Small Latex cleanup.

@@
   \begin{array}[t]{l}
-    \text{$\dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand$} \\
-    \text{$\dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand$} \\
-    \text{$\dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand$} \\
+    \dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand \\
+    \dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand \\
+    \dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand \\
     \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
-      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r) \nsubsim \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsubeq \stdreg{R(r),\gc}{\pur}
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsubsim \stdreg{R(r),\gc}{\pur}
     \end{array}
@@
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
-      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \quad \forall r \in S \ldotp  \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
+      \qquad \forall r \in S \ldotp  \pwheap(r) \nsim \codereg{\_,\_,M(r),\gc}
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad \lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1\\
-      \qquad \quad \forall r \in S \ldotp \pwheap(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r) \nsupsim \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \quad \forall r \in S \ldotp |R(r)| = 1 \\
-      \qquad \quad \forall r \in S \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsupsim \stdreg{R(r),\gc}{\pur} \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}
@@
     \item $\trust = \trusted$
-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
+    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}\updReg{\pcreg}{((\rx,\lin),\baddr,\eaddr,\aaddr)}, \src{\stk}, \src{\ms_\stk})$ and
       \begin{itemize}
@@
       \end{itemize}
-      \dominique{something missing here: pc?}
     \end{itemize}

### 2018-06-06 fully-abstract-comp/technical-report.tex
Component well-formedness no longer preserves linking.

--- 2018-06-05 fully-abstract-comp/technical-report.tex
@@
     \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
-    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}}
+    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta)
   }{
@@
     A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
-    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]\\
+    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
   } {
@@
 
-\begin{lemma}[Linking preserves component well-formedness]
-  \label{lem:linking-well-formedness}
-  If $\vdash \var{comp}_1$ and $\vdash \var{comp}_2$, then $\vdash \var{comp}_1 \bowtie \var{comp}_2$.
-\end{lemma}
-\begin{proof}
-  Follows by induction on the component validity judgement and definition-chasing.
-  The most interesting point is related to imports that are resolved from the other component during linking, but we know that those exports satisfy the $\mathrm{comp-value}$ judgement.
-\end{proof}
+% update: linking no longer preserves well-formedness because we require is entirely trusted or not trusted at all.
+% \begin{lemma}[Linking preserves component well-formedness]
+%   \label{lem:linking-well-formedness}
+%   If $\vdash \var{comp}_1$ and $\vdash \var{comp}_2$, then $\vdash \var{comp}_1 \bowtie \var{comp}_2$.
+% \end{lemma}
+% \begin{proof}
+%   Follows by induction on the component validity judgement and definition-chasing.
+%   The most interesting point is related to imports that are resolved from the other component during linking, but we know that those exports satisfy the $\mathrm{comp-value}$ judgement.
+% \end{proof}
 

### 2018-06-06 fully-abstract-comp/technical-report.tex
Add a label.
Add a new notion of disjoint sum for worlds as memory presence specifications rather than authority specifications.

@@
 \begin{lemma}
+  \label{lem:oplus-assoc-comm}
   $\oplus$ is associative and commutative.
@@
   Follows easily from the definitions.
+\end{proof}
+
+We also define a second disjoint union operator of worlds:
+\[
+  W_1 \uplus W_2 = W
+  \text{ iff }
+  \begin{array}[t]{l}
+    \dom(\pwheap) = \dom(\pwheap[W_1]) \uplus \dom(\pwheap[W_2]) \tand \\
+    \dom(\pwfree) = \dom(\pwfree[W_1]) \uplus \dom(\pwfree[W_2]) \tand \\
+    \dom(\pwpriv) = \dom(\pwpriv[W_1]) \uplus \dom(\pwpriv[W_2]) \\
+  \end{array}
+\]
+The two operators $\uplus$ and $\oplus$ are quite different.
+The difference is most clear in the treatment of pure regions: $\uplus$ allows both worlds to have the same pure region, while $\oplus$ forbids this.
+To understand this different treatment ($W_1 \uplus W_2$ and $W_1 \oplus W_2$), you should understand that the two are intended for different usages of worlds.
+The $W_1 \oplus W_2$ operator treats the worlds as specifications of authority: taking the disjoint union of worlds specifying non-exclusive ownership of a block of memory is allowed and produces a new world that also specifies non-exclusive ownership of world.
+The $W_1 \uplus W_2$  operator treats worlds as specifications of memory contents: taking the disjoint union of worlds specifying the presence of the same memory range is not allowed.
+The latter operator is used in the logical relation for components which specifies (among other things) that the world should specify the presence of the component's data memory.
+Linking two components then produces a new component with both components' data memory.
+The linked component is then valid in a world that has the combined memory presence specifications, not the combined authority.
+In other words, $\oplus$ specifies disjoint authority distribution, while $\uplus$ specifies disjoint memory allocation.
+
+Note also that this picture is further complicated by our usage of non-authority-carrying $\spatial$ regions.
+They are really only there in a world $W$ as a shadow copy of a $\spatialo$ region in another world $W'$ that $W$ will be combined with.
+The shadow copy is used for specifying when a memory satisfies a world: the memory should contain all memory ranges that anyone has authority over, not just the ones whose authority belongs to the memory itself.
+For example, if a register contains a linear pointer to a range of memory, then the register file will be valid in a world where the corresponding region is $\spatialo$, while the memory will be valid in a world with the corresponding region only $\spatial$.
+However, for the memory to satisfy the world, the block of memory needs to be there, i.e.
+the memory should contain blocks of memory satisfying every region that is $\spatialo$, $\pure$, but also just $\spatial$ (because it may be $\spatialo$ in, for example, the register file's world).
+
+\begin{lemma}
+  \label{lem:uplus-assoc-comm}
+  $\uplus$ is associative and commutative.
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
+
+\begin{lemma}[Odd distributivity of $\oplus$ and $\uplus$]
+  \label{lem:oplus-distr-uplus}
+  \begin{equation*}
+    (W_1\oplus W_2) \uplus (W_3 \oplus W_4) = (W_1 \uplus W_3) \oplus (W_2 \uplus W_4)
+  \end{equation*}
+  Also, the left expression is defined iff the right expression is.
+\end{lemma}
+\begin{proof}
+  Follows by definition-chasing.
 \end{proof}
@@
 
+\begin{lemma}[Adding additional memory is the future]
+  \label{lem:uplus-future}
+  For all $W_1,W_2$, we have that:
+  $W_1 \uplus W_2 \future W_1$
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
+
 \begin{lemma}[Purity is a thing of the past]
@@
 
+\begin{lemma}[Combined independent heap memory satisfies disjoint world]
+  \label{sec:combined-memory-disjoint-world}
+  If $\memSatHeap{\overline{\sigma_1},ms_{S,1},\ms_{T,1}}{W_1}$ and $\memSatHeap{\overline{\sigma_2},ms_{S,2},\ms_{T,2}}{W_2}$, then $\memSatHeap{\overline{\sigma_1}\uplus\overline{\sigma_2},ms_{S,1}\uplus \ms_{S,2},\ms_{T,1}\uplus \ms_{T,2}}{W_1 \uplus W_2}$.
+\end{lemma}
+\begin{proof}
+  Unfolding the definitions, it's easy to construct the memory and seal partitions $R_{\ms,3}$ and $R_{\var{seal},3}$ from the corresponding partitions of the separate memories.
+  Constructing $R_{W,3}$ from $R_{W,1}$ and $R_{w,2}$ is a bit harder, but we can use Lemmas~\ref{lem:oplus-distr-uplus},~\ref{lem:purePart-duplicable},~\ref{lem:uplus-assoc-comm} and~\ref{lem:oplus-assoc-comm} to manipulate the world equalities as follows:
+  \begin{align*}
+    W_1 \uplus W_2 &= (\purePart{W_1} \oplus W_1) \uplus (W_2 \oplus \purePart{W_2})\\
+                   &= (\purePart{W_1} \uplus W_2) \oplus (W_1 \uplus \purePart{W_2})\\
+                   &= (\purePart{W_1} \uplus \bigoplus_r R_{W,2}(r)) \oplus ((\bigoplus_r R_{W,1}(r)) \uplus \purePart{W_2})\\
+                   &= \bigoplus_r (\purePart{W_1} \uplus R_{W,2}(r)) \oplus \bigoplus_r (R_{W,1}(r) \uplus \purePart{W_2})
+  \end{align*}
+  We can then define
+  \begin{equation*}
+    R_{W,3}(r) =
+    \begin{cases}
+      \purePart(W_1) \uplus R_{W,2}(r) &\text{when } r \in \dom(R_{W,2})\\
+      \purePart(W_2) \uplus R_{W,1}(r) &\text{when } r \in \dom(R_{W,1})
+    \end{cases}
+  \end{equation*}
+  Lemma~\ref{lem:uplus-future} and the monotonicity of the regions' memory predicates then yield the required results.
+\end{proof}
+
 \subsection{Relation}
@@
   \begin{itemize}
-  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W)$.
+  \item $\npair{(\var{comp}_1 \bowtie \var{comp}_2,\var{comp}_1\bowtie\var{comp}_2)} \in \lrcomp(W_1 \uplus W_2)$.
   \end{itemize}

### 2018-06-06 fully-abstract-comp/technical-report.tex
Answer for one of Dominique's comments.

@@
   This also means we can afford to use ``up to n'' without breaking contractiveness of the recursive definition of the value relation.
-  }
-We have the following notion of regions being an $n$-subset of another region.
+  }\\
+\lau{You cannot take $\nsubsim$ on a UPred. It only works on regions. We use the
+  special notation to signify that it is almost like standard $n$-equality. I suppose
+  you are thinking of $\nequal$ for UPreds? The way I have seen it before, it is defined as
+``up to $n-1$ (including)''. More precisely: Define the $k$-cut of $p \in \UPred{A}$ to be
+\newcommand{\floor}[1]{\lfloor #1 \rfloor}
+\newcommand{\kcut}[2][k]{\floor{#2}_{#1}}
+  \[
+  \kcut{p} = \{ (j,a) \in p | j < k \}.
+  \]
+  We use this to define a family of equivalences $(\nequal)_{n=0}^{\infty}$ to be, for $p$ and $q$ in $\UPred{A}$
+  \[
+    p \nequal q \textit{ iff } \kcut[n]{p} = \kcut[n]{q}
+  \]
+  If we pick a different definition, we should check that it is indeed a
+  well-defined $n$-equality (for instance, it should be the total relation for $\nequal[0]$).
+} We have the following notion of regions being an $n$-subset of another region.
 \begin{mathpar}

### 2018-06-06 fully-abstract-comp/technical-report.tex
Added condition to call, so seal must be within bounds of pc.
Useful lemmas

@@
       \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
-      \mem(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
+      \mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
+      \baddr \leq  \aaddr+\offpc \leq \eaddr \tand \\
       \sigma_? = \sigma_\aaddr + \offsigma \tand\\
@@
 \section{Lemmas}
+\begin{lemma}
+  \label{lem:unique-h-sigma}
+  if $\iota,iota' \in \dom(\pwheap)$ and $\pwheap(\iota) = (\pure,\_,H_\sigma)$
+  and $\pwheap(\iota') = (\pure,\_,H_\sigma')$ and $\sigma \in \dom(H_\sigma)$
+  and $\sigma \in \dom(H_\sigma')$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$,
+  then
+  \[
+    H_\sigma = H_\sigma' \text{ and } \iota = \iota'
+  \]
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
+
+\begin{lemma}
+  if $\npair{[\baddr,\eaddr]} \in \stackReadCond{W}$, and
+  $\npair{[\baddr,\eaddr]} \in \stackWriteCond{W}$, and\\
+  $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, then
+  \[
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\linear,\pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp \\
+      \qquad \biguplus R(r) = [\baddr,\eaddr] \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r) \nsim \stdreg{R(r),\gc}{\pur} \wedge \\
+      \qquad \quad |R(r)| = 1 \wedge \\
+      \qquad \quad \pwfree(r) \text{ is address-stratified}
+    \end{array}
+  \]
+\end{lemma}
+\begin{proof}
+  By assumption we get
+  \begin{enumproof}
+    \item $S_R \subseteq \addressable{\linear,\pwfree}$
+    \item $R_R : S_R \fun \powerset{\nats}$
+    \item $S_W \subseteq \addressable{\linear,\pwfree}$
+    \item $R_W : S_W \fun \powerset{\nats}$
+  \end{enumproof}
+  such that
+  \begin{enumproof}[resume]
+    \item $\biguplus R_R(r) = [\baddr,\eaddr]$
+    \item $\forall r \in S_R \ldotp \pwfree(r) \nsubsim
+      \stdreg{R_R(r),\gc}{\pur} \text{ and } |R_R(r)|=1$
+    \item $\biguplus R_W(r) = [\baddr,\eaddr]$
+    \item $\forall r \in S_W \ldotp \pwfree(r) \nsupsim
+      \stdreg{R_W(r),\gc}{\pur} \text{ and } |R_R(r)|=1 \text{ and } \pwfree(r)
+      \text{ is address-stratified}$
+  \end{enumproof}
+  Now pick $S = S_R \cup S_W$ and show
+  \[
+    \forall r \in S \ldotp R_R(r) = R_W(r)
+  \]
+  Assuming for contradiction $\exists r \in S \ldotp R_R(r) \neq R_W(r)$, we have $\pwfree(r) \nsubsim \stdreg{*,\gc}{\pur}$ and $\pwfree(r) \nsupsim \stdreg{-,\gc}{\pur}$ for two distinct singleton sets $*$ and $-$.
+  Due to the memory satisfaction assumption $\pwfree(r)$ is non-empty and it is trivial to show that so is the standard regions.
+  From the above, we can conclude that $\pwfree(r)$ should contain some $\ms$
+  with $\dom(\ms) = *$ and $\dom(\ms) = -$ which contradicts the address
+  stratification assumption.
+
+  Now pick
+  \[
+    R = R_R(r) \text{ for $r \in S$}
+  \]
+  We first show $\biguplus R(r) \supseteq [\baddr,\eaddr]$.
+  For $\aaddr \in [\baddr,\eaddr]$ we know that there exists $r_1$ and $r_2$ such that $R_W(r_1) = R_R(r_2) = \{\aaddr\}$.
+  It suffices to show $r_1 = r_2$.
+  To this end assume the contrary.
+  This, the address-stratification assumption on all write regions, the $n$-subset of a standard region for read regions, and memory satisfaction lead to a contradiction as any part of memory only can be governed by one region.
+  Address-stratification and the singleton requirement follows trivially from
+  the assumptions from the write region. Finally the $n$-equality follows from Lemma~\ref{lem:nsub-and-nsup-std}.
+\end{proof}
 

### 2018-06-06 fully-abstract-comp/ftlr-proof-cases.tex
FTLR proof changes

@@
     \item $\exec{\Phi_S(\pcreg)}$
-    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in
-      \ints$.
+    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in \ints$.
+    \item $[\baddr,\eaddr] \subseteq \ta$
     \item $\Phi_S \step[\gc] \Phi_S' \neq \failed$
@@
     \item $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$
-    \item
-      \[
+    \item \[
         \begin{split}
-          \Phi_S'.\reg = \Phi_S.\reg
-          &[\pcreg \mapsto c_1] \\
-          &[\rdata \mapsto c_2] \\
-          &[\rstk \mapsto \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}] \\
-          &[\rretc \mapsto \sealed{\sigma',\retptrc{\baddr,\eaddr,\aaddr+\calllen}}]\\
-          &[\rretd \mapsto \sealed{\sigma',\retptrd{\aaddr_\stk,\eaddr_\stk}}] \\
+          \Phi_S''.\reg = \Phi_S.reg      &[\rstk \mapsto \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}] \\
+          &[\rretc \mapsto \sealed{\sigma',\retptrc(\baddr,\eaddr,\aaddr+\calllen)}]\\
+          &[\rretd \mapsto \sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)}] \\
           &[r_1,r_2 \mapsto w_1,w_2] \\
@@
       \]
-    \item $\Phi_S'.\mem = \Phi_S.\mem$
-    \item $\Phi_S'.\stk = ((\aaddr+\calllen),\ms_\stk |_{\aaddr_\stk,\eaddr_\stk}\update{\aaddr_\stk}{42}) :: \Phi_S.\stk$
-    \item $\Phi_S'.\ms_\stk = \Phi_S.\ms_\stk = \ms_\stk - \ms_\stk|_{\aaddr_\stk,\eaddr_\stk}$
+    \item $\Phi_S''.\mem = \Phi_S.\mem$
+    \item $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{\aaddr_\stk,\eaddr_\stk}\update{\aaddr_\stk}{42}$
+    \item $\Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk$
+    \item $\Phi_S''.\ms_\stk = \Phi_S.\ms_\stk -
+      \Phi_S.\ms_\stk|_{\aaddr_\stk,\eaddr_\stk}$
+    \item $\baddr \leq \aaddr+\offpc \leq \eaddr$
+      \lau{TODO: Need to have an explicit $\offpc$.}
     \item $\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}$
@@
     \item $\sigma_\aaddr \leq \sigma' \leq \sigma_\eaddr$
+
+    \item $\Phi_S' =\var{xjumpResult}\left(c_1,c_2, \Phi_S' \right)$
+
     \end{itemize}
@@
 as desired.
+\\\\
+In case \ref{case:ftlr:call}, we first need to let the target execution catch
+up. That is $\Phi_T' \nstep[15]{} \Phi_T''$ for 
+\[
+\Phi_T'' = \Phi_T
+\begin{array}[t]{l}
+  \update{\mem.\aaddr_\stk}{42} \\
+  \updReg{\rstk}{((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1)} \\
+  \updReg{\rretd}{\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}} \\
+  \updReg{\rretc}{\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\var{off\_return})}} \\
+  \updReg{\rtmp{1}}{0}\\
+  \updReg{r_1,r_2}{\linCons{\Phi_T(r_1)},\linCons{\Phi_T(r_2)}}\\
+  \updReg{\pcreg}{c_1'}\\
+  \updReg{\rdata}{c_2'}
+\end{array}
+\]
+where $\Phi_T(r_i) = \sealed{\sigma,c_i'} \text{ for $i\in\{1,2\}$}$ and $\var{off\_return} = 15$ which is the offset to the return code.
+Now using Lemma~\ref{lem:lro-anti-red-gen} again, it suffices to show
+\[
+  \npair[n'-1]{(\Phi_S',\Phi_T'')} \in \lro
+\]
+By assumption, we have $\npair[n']{(\reg_S(r_i),\reg_T(r_i))} \in \lrr(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
+We know the capabilities in $r_1$ and $r_2$ are sealed capabilities, and by
+Lemma~\ref{lem:unique-h-sigma} and the definition of $\lrv$ we get
+$\npair[n'-1]{\stpair[.]{c_2}{c_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$
+and w.l.o.g.\
+
+\begin{equation}
+\forall W' \future W_{R,1}, W_o, n' < n, \npair[n']{(c_2,c_2')} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \npair[n']{c_1,c_2,c_1',c_2'} \in \lrexj(W'\oplus W_o)\label{eq:ftlr:call:sealed-code-assump}
+\end{equation}
+Now construct $W_{R,1}'$ as follows:
+
+By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r$ in $S$ $\pwfree[W_{R,1}](r) \nsim \stdreg{R(r),\gc}{\pur}$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
+Now define
+\[
+  W_{R,1}' = W_{R,1}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto \stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square}]
+\] 
+For a fresh $r_\var{priv\_stk}$.
+We know $W_{R,1}' \future W_{R,1}$ as the revoked regions must have been $\spatial$ in $W_{R,q}$ (as they are owned by the part of the world assigned to the stack-register in the register-file relation).
+The static region for the private stack is an extension.
+
+Pick this world as $W'$ in Eq~\ref{eq:ftlr:call:sealed-code-assump}
+. Let $W_{R,2}'$ be the same world but
+with the ownership of $W_{R,2}$ and pick it for $W_o$.
+Now use monotonicity of $H_\sigma$ with the above facts to get
+\[
+  \npair[n']{c_1,c_2,c_1',c_2'} \in \lrexj(W_{R,1}'\oplus W_{R,2}')
+\]
+Now pick register files and memories such that they form $\Phi_S''$ (defined in
+the assumptions) and $\Phi_T''$ and for $W_R'$ and $W_M'$ (we define them below) show
+\begin{enumerate}[label=\roman*.]
+\item \label{case:ftlr:scall:world} $W_{R,1}'\oplus W_{R,2}' \oplus W_R' \oplus W_M'$
+\item \label{case:ftlr:scall:reg} $\npair[n'-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$
+\item \label{case:ftlr:scall:mem} $\memSat[n'-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}$
+\end{enumerate}
+to get
+\[
+ \npair[n'-1]{(\Phi_S',\Phi_T'')} \in \lro
+\]
+as desired. It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}.
+
+Case \ref{case:ftlr:scall:world}: Pick $W_R'$ and $W_M'$ to have the regions of $W_{R,1}'$, but where $W_R'$ owns $r_\var{priv\_stk}$ and otherwise has the ownership of $W_R$ and $W$ with the exception of the regions owned by $W_{R,1}'$ and $W_{R,2}'$.
+$W_M'$ has the ownership of $W_M$.
+Case \ref{case:ftlr:scall:world} follows from assumption $W \oplus W_R \oplus W_M$.
+The only changes to the worlds is that some ownership has been shifted from $W_R$ to $W_{R,1}$ and $W_{R,2}$ and the ownership for $W$ now belongs to $W_R'$.
+In other words, no ownership has been duplicated.
+
+Case \ref{case:ftlr:scall:reg}, using assumption $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
+We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership.
+region $\rstk$ maps to a world with the same ownership, but of course without the now revoked regions.
+Region $\rretd$ maps to a world that owns private $r_\var{priv\_stk}$ region.
+Finally, $\rretc$ maps to a world with the ownership of $W$.
+
+we split the world in the same way for the registers that remain unchanged, the
+result follows from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity}.
+This leaves the following cases:
+\begin{description}
+\item[Case $\rstk$:] \lau{TODO}
+\item[Case $\rtmp{1}$:] \lau{TODO}
+\item[Case $r_1$,$r_2$:] \lau{TODO}
+\item[Case $\rretd$:] We have to show
+  \[
+    \npair[n'-1]{(\sealed{\sigma',\retptrc(\aaddr_\stk,\eaddr_\stk)},\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk)})} \in \lrv(W_{R,\rretd}')
+  \]
+  where $W_{R,\rretd}'$ is the part of $W_R'$ with ownership over $r_\var{priv\_stk}$.
+  From assumption $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$\footnote{This must be the other possible assumption contains $[\baddr,\eaddr] \# \ta$ which leads to a contradition with $[\baddr,\eaddr] \subseteq \ta$}, we know there is a code-region for the
+
+\item[Case $\rretc$:] \lau{TODO}
+\end{description}
+
+\begin{lemma}
+  If $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss,\_)$ and
+  $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and
+  $\exists \aaddr \ldotp \mscode(\aaddr) = \sigma$ and
+  $\sigma \in \sigrets$ and
+  $\memSat{\dots}{W}$, then
+  \[
+    \begin{array}{l}
+      \exists \mscode' r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) = \codereg{\sigrets',\sigcloss',\mscode',(\ta,\stkb)} \text{ and } \\
+      \quad [\baddr,\eaddr] \subseteq \dom(\mscode') \text{ and } \\
+      \quad \sigma \in \sigrets' \subseteq \sigrets \text{ and } \\
+      \quad \sigclos \subseteq \sigclos'
+    \end{array}
+  \]
+\end{lemma}
 

### 2018-06-06 fully-abstract-comp/technical-report.tex
Typos

@@
     \begin{cases}
-      \purePart(W_1) \uplus R_{W,2}(r) &\text{when } r \in \dom(R_{W,2})\\
-      \purePart(W_2) \uplus R_{W,1}(r) &\text{when } r \in \dom(R_{W,1})
+      \purePart{W_1} \uplus R_{W,2}(r) &\text{when } r \in \dom(R_{W,2})\\
+      \purePart{W_2} \uplus R_{W,1}(r) &\text{when } r \in \dom(R_{W,1})
     \end{cases}
@@
     \begin{array}{l}
-      \exists S \subseteq \addressable{\lin, \pwfree} \ldotp \\
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
       \quad \exists M : S \fun \powerset{\MemSeg} \\
@@
   \begin{equation*}
-    H^\mathrm{code}\ ;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubsim H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
+    H^\mathrm{code}\;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubsim H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
   \end{equation*}
-  So take $\npair{\ms_S,\ms_T} \in H^\mathrm{code}\ ;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W}$, then we know that
+  So take $\npair{\ms_S,\ms_T} \in H^\mathrm{code}\;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W}$, then we know that
   \begin{itemize}
@@
 \begin{proof}
-  Follows by defintion of $\readCond$ from Lemma~\ref{lem:codereg-untrusted-stdreg}.
+  Follows by defintion of $\readCond{}$ from Lemma~\ref{lem:codereg-untrusted-stdreg}.
 \end{proof}

### 2018-06-07 fully-abstract-comp/ftlr-proof-cases.tex
Typos
Removed nsubsim relation and updated documents accordingly. xReadCondition now defaults to normal.

@@
   \item If $\npair{A} \in \execCond{\lin,W}$, then $\npair[n']{A} \in \execCond{\lin,W}$.
-  \item If $\npair{A} \in \xReadCond{\lin,W}$, then $\npair[n']{A} \in \xReadCond{\lin,W}$.
+  \item If $\npair{A} \in \xReadCond{W}$, then $\npair[n']{A} \in \xReadCond{W}$.
   \item If $\npair{(w,w')} \in \lrvg{\trust}{W}$, then $\npair[n']{(w,w')} \in \lrvg{\trust}{W}$.
@@
   \begin{itemize}
-  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \nequal \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W'}$.
+  \item $\xReadCond{W} \nequal \xReadCond{W'}$.
   \item ...
@@
   From $\npair{(b,e)}\in\readCond{l,W}$, we get an $S \subseteq \addressable{\lin,
-    \pwheap}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwheap(r)
-  \nsubsim \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
+    \pwheap}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwheap(r).H
+  \nsubeq \stdreg{R(r),\gc}{\pur}.H$ for all $r \in S$.
 
@@
   \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
-  Because $\pwheap(r) \nsubsim \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
-  also $\pwheap[W_\var{heap}](r) \nsubsim \stdreg{R(r),\gc}{\pur}$. This
+  Because $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
+  also $\pwheap[W_\var{heap}](r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$. This
   means that also $\npair{R_\ms(r)} \in
-  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined
-    $\nsubsim$ somewhere?} \lau{Yes, it is in subsection~\ref{subsec:worlds}}
+  H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.
 
@@
   we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
-  for all $r \in S'$, $\pwheap(r) \nsupsim \stdreg{R'(r),\gc}{\pur}$ and
+  for all $r \in S'$, $\pwheap(r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$ and
   $\pwheap(r) \text{ is address-stratified}$.
@@
   Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwheap[W_\var{heap}](r') \nsupsim \stdreg{R'(r'),\gc}{\pur}$ and
+  $\pwheap[W_\var{heap}](r').H \nsupeq \stdreg{R'(r'),\gc}{\pur}.H$ and
   $\pwheap[W_\var{heap}](r') \text{ is address-stratified}$.
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r) \nsupsim \stdreg{R'(r),\gc}{\pur}$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
 
@@
   From $\npair{(b,e)}\in\stackReadCond{W}$, we get an $S \subseteq \addressable{\lin,
-    \pwfree}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwfree(r)
-  \nsubsim \stdreg{R(r),\gc}{\pur}$ for all $r \in S$.
+    \pwfree}$, an $R : S \rightarrow \powerset{\nats}$ with $\biguplus_{r \in S} R(r) \supseteq [\baddr,\eaddr]$ and $\pwfree(r).H
+  \nsubeq \stdreg{R(r),\gc}{\pur}.H$ for all $r \in S$.
 
@@
   We have that $r \in \addressable{\lin, \pwfree} \subseteq \activeReg{\pwfree[W_\var{free\_stack}]}$, so $\npair{R_\ms(r)} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
-  Because $\pwfree(r) \nsubsim \stdreg{R(r),\gc}{\pur}$ and $W\oplus W_M = W \oplus (W_\var{free\_stack} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that also $\pwfree[W_\var{free\_stack}](r) \nsubsim \stdreg{R(r),\gc}{\pur}$.
-  This means that also $\npair{R_\ms(r)} \in H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.\dominique{have we defined $\nsubsim$ somewhere?}
-  \lau{Yes, it is in subsection~\ref{subsec:worlds}}
+  Because $\pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$ and $W\oplus W_M = W \oplus (W_\var{free\_stack} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that also $\pwfree[W_\var{free\_stack}](r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$.
+  This means that also $\npair{R_\ms(r)} \in H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.
 
@@
   we get an $S'\subseteq \addressable{\lin, \pwfree} \subseteq \activeReg{\pwheap[W_\var{free\_stack}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ and
-  for all $r \in S'$, $\pwfree(r) \nsupsim \stdreg{R'(r),\gc}{\pur}$ and
+  for all $r \in S'$, $\pwfree(r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$ and
   $\pwfree(r) \text{ is address-stratified}$.
@@
   Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwfree[W_\var{free\_stack}](r') \nsupsim \stdreg{R'(r'),\gc}{\pur}$ and
+  $\pwfree[W_\var{free\_stack}](r').H \nsupeq \stdreg{R'(r'),\gc}{\pur}.H$ and
   $\pwfree[W_\var{free\_stack}](r') \text{ is address-stratified}$.
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r) \nsupsim \stdreg{R'(r),\gc}{\pur}$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
 
@@
 
-  It remains to prove that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
+  it remains to prove that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
   We have to distinguish the case that $\isLinear{\ms_S(\aaddr)}$ and the opposite case.
@@
   \begin{itemize}
-  \item $\trust = \trusted \vee \npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$
-  \item $\npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W}$
+  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
   \end{itemize}
@@
 
-By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r$ in $S$ $\pwfree[W_{R,1}](r) \nsim \stdreg{R(r),\gc}{\pur}$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
+By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r$ in $S$ $\pwfree[W_{R,1}](r).H \nequal \stdreg{R(r),\gc}{\pur}.H$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
 Now define
@@
   where $W_{R,\rretd}'$ is the part of $W_R'$ with ownership over $r_\var{priv\_stk}$.
-  From assumption $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$\footnote{This must be the other possible assumption contains $[\baddr,\eaddr] \# \ta$ which leads to a contradition with $[\baddr,\eaddr] \subseteq \ta$}, we know there is a code-region for the
+  From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get a
+  set of regions $R \subseteq \addressable{\normal, \pwheap}$ 
+and
+  $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W}$,
 
@@
 \begin{lemma}
-  If $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss,\_)$ and
+
+  If $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ and
+  $\memSat{\ms_S \uplus \mscode, \stk,\ms_\stk,\ms_T}{W}$
   $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and
-  $\exists \aaddr \ldotp \mscode(\aaddr) = \sigma$ and
-  $\sigma \in \sigrets$ and
-  $\memSat{\dots}{W}$, then
   \[
     \begin{array}{l}
-      \exists \mscode' r \in \dom(\pwheap) \ldotp \\
+      \exists r \in \dom(\pwheap) ,\mscode' \ldotp \\
       \quad \pwheap(r) = \codereg{\sigrets',\sigcloss',\mscode',(\ta,\stkb)} \text{ and } \\

### 2018-06-07 fully-abstract-comp/technical-report.tex
no hidden calls requirement added to code memory judgement.
Correction to "no hidden calls" definition.
The first attempt allowed for no call. Now it requires either a call or some
value in memory that shows that it is not a call.
Removed nsubsim relation and updated documents accordingly. xReadCondition now defaults to normal.

@@
   \inference{
+    \text{$\mscode$ has no hidden calls}\\
     \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
@@
 \end{mathpar}
+where the following definition is used
+\begin{definition}[No hidden calls]
+  \label{def:no-hidden-calls}
+  We say that a memory segment $\mscode$ has no hidden calls iff
+  \[
+    \begin{array}{l}
+      \forall \aaddr \in \dom(\mscode) \ldotp \\
+      \quad \forall i \in [0,\calllen - 1] \\
+      \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
+      \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1]
+      \wedge \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
+      \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
+      \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
+    \end{array}
+  \]
+\end{definition}
 
@@
 
-\dominique{Remaining question: what exactly does $\nsubsim$ on a UPred mean?  Does it mean subset ``up to n (including)'' or ``up to n-1 (including)''?
-  Check what we can require without breaking contractiveness of recursive definition of value relation and/or what we did in previous paper and/or what we need in FTLR proof.
-  Update: in the FTLR proof, we seem to need ``up to n'' because the heap predicate of the standard region is contractive in terms of the value relation.
-  This also means we can afford to use ``up to n'' without breaking contractiveness of the recursive definition of the value relation.
-  }\\
-\lau{You cannot take $\nsubsim$ on a UPred. It only works on regions. We use the
-  special notation to signify that it is almost like standard $n$-equality. I suppose
-  you are thinking of $\nequal$ for UPreds? The way I have seen it before, it is defined as
-``up to $n-1$ (including)''. More precisely: Define the $k$-cut of $p \in \UPred{A}$ to be
-\newcommand{\floor}[1]{\lfloor #1 \rfloor}
-\newcommand{\kcut}[2][k]{\floor{#2}_{#1}}
+% \dominique{Remaining question: what exactly does $\nsubsim$ on a UPred mean?  Does it mean subset ``up to n (including)'' or ``up to n-1 (including)''?
+%   Check what we can require without breaking contractiveness of recursive definition of value relation and/or what we did in previous paper and/or what we need in FTLR proof.
+%   Update: in the FTLR proof, we seem to need ``up to n'' because the heap predicate of the standard region is contractive in terms of the value relation.
+%   This also means we can afford to use ``up to n'' without breaking contractiveness of the recursive definition of the value relation.
+%   }\\
+\lau{I suppose you are thinking of $\nequal$ for UPreds?
+  The way I have seen it before, it is defined as ``up to $n-1$ (including)''.
+  More precisely: Define the $k$-cut of $p \in \UPred{A}$ to be \newcommand{\floor}[1]{\lfloor #1 \rfloor} \newcommand{\kcut}[2][k]{\floor{#2}_{#1}}
   \[
@@
   \]
-  If we pick a different definition, we should check that it is indeed a
-  well-defined $n$-equality (for instance, it should be the total relation for $\nequal[0]$).
-} We have the following notion of regions being an $n$-subset of another region.
-\begin{mathpar}
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-            { (v,H) \nsubsim (v',H') }
-\and
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-            { (\pure,H,H_\sigma) \nsubsim (v',H') }
-\and
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-            { (v,H) \nsubsim (\pure,H',H_\sigma) }
-\and
-  \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-            { (\pure,H,H_\sigma) \nsubsim (\pure,H',H_\sigma') }
-\end{mathpar}
-We define $n$-equality of regions $\iota \nsim \iota'$ as the symmetric closure, i.e. $\iota \nequal \iota'$ iff $\iota \nsubsim \iota'$ and $\iota \nsupsim \iota'$.
+  If we pick a different definition, we should check that it is indeed a well-defined $n$-equality (for instance, it should be the total relation for $\nequal[0]$).
+} % We have the following notion of regions being an $n$-subset of another region.
+% \begin{mathpar}
+%   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+%             { (v,H) \nsubsim (v',H') }
+% \and
+%   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+%             { (\pure,H,H_\sigma) \nsubsim (v',H') }
+% \and
+%   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+%             { (v,H) \nsubsim (\pure,H',H_\sigma) }
+% \and
+%   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
+%             { (\pure,H,H_\sigma) \nsubsim (\pure,H',H_\sigma') }
+% \end{mathpar}
+% We define $n$-equality of regions $\iota \nsim \iota'$ as the symmetric closure, i.e. $\iota \nequal \iota'$ iff $\iota \nsubsim \iota'$ and $\iota \nsupsim \iota'$.
 
@@
         \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{\lin,W} \wedge\\
-                                  \npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}\endarray
+                                  \npair{[\baddr,\eaddr]} \in \xReadCond{W}\endarray
       \end{array}
@@
       \perm \sqsubseteq \rx \wedge \\
-      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,(\ta,\stkb)]{\normal,W}
+      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,(\ta,\stkb)]{W}
     \end{array}
@@
       \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nsubsim \stdreg{R(r),\gc}{\pur}
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
     \end{array}
@@
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsubsim \stdreg{R(r),\gc}{\pur}
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
     \end{array}
@@
 \[
-  \xReadCond{\lin,W} = \left\{ \npair{A} \middle| 
-    \begin{array}{l}
-      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+  \xReadCond{W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\normal, \pwheap} \ldotp \\
       \quad \exists M : S \fun \powerset{\MemSeg} \\
       \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
-      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
       % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \forall r \in S \ldotp  \pwheap(r) \nsim \codereg{\_,\_,M(r),\gc}
+      \qquad \forall r \in S \ldotp \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
     \end{array}
@@
       \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nsupsim \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
       \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsupsim \stdreg{R(r),\gc}{\pur} \wedge\\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}
@@
 \]
-\lau{02-11-2017: I am not sure whether it should be a $\normal$ region. In previous work $\nsubsim$ ignored the ``view'' which it may also do here, and then then it does not matter which one it is.}
 where $H^{\mathrm{std},\square}_A$ is defined as follows:
@@
         \dom(\code) \subseteq \ta \tand\\
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r) \nsim (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
@@
   \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
-  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W}$
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
   \end{itemize}
@@
   \end{itemize}
-  then $\codereg{\sigrets,\sigcloss,\mscode,\gc} \nsubsim \stdreg{\dom(\mscode),\gc}{\pur}$ 
+  then $\codereg{\sigrets,\sigcloss,\mscode,\gc}.H \nsubeq \stdreg{\dom(\mscode),\gc}{\pur}.H$ 
 \end{lemma}
@@
   \begin{equation*}
-    H^\mathrm{code}\;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubsim H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
+    H^\mathrm{code}\;\sigrets\;\sigcloss\;\mscode\;\gc \; \hat{W} \nsubeq H_{\dom(\mscode)}^{\mathrm{std},\square}\;\hat{W}
   \end{equation*}
@@
     $\npair{A'} \in \writeCond{\lin,W}$.
-  \item If $\npair{A} \in \xReadCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
-    $\npair{A'} \in \xReadCond{\lin,W}$.
+  \item If $\npair{A} \in \xReadCond{W}$  and $\emptyset \neq A' \subseteq A$, then
+    $\npair{A'} \in \xReadCond{W}$.
   \item If $\npair{A} \in \execCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
@@
   \begin{itemize}
-  \item $\npair{\dom(\mscode)}\in\xReadCond{\normal,W}$
+  \item $\npair{\dom(\mscode)}\in\xReadCond{W}$
   \item ($\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$) or ($\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$)
@@
       this follows from Lemma~\ref{lem:code-memory-untrusted-readable}.
-    \item $\npair{[b,e]}\in\xReadCond{\normal,W}$:
+    \item $\npair{[b,e]}\in\xReadCond{W}$:
       by assumption and Lemma~\ref{lem:conds-shrinkable}.
@@
   \item $\trust = \trusted$ and $\dom(\mscode) \subseteq \ta$:
-    By definition of $\lrvg{\trusted}(W)$ and Lemma~\ref{lem:conds-shrinkable} with the assumption that $\npair{\dom(\mscode)}\in\xReadCond{\normal,W}$.
+    By definition of $\lrvg{\trusted}(W)$ and Lemma~\ref{lem:conds-shrinkable} with the assumption that $\npair{\dom(\mscode)}\in\xReadCond{W}$.
   \end{itemize}
@@
 
-     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs} that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(W)$, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs} that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(W)$, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
 
@@
       \qquad \biguplus R(r) = [\baddr,\eaddr] \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r) \nsim \stdreg{R(r),\gc}{\pur} \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nequal \stdreg{R(r),\gc}{\pur}.H \wedge \\
       \qquad \quad |R(r)| = 1 \wedge \\
@@
     \item $\biguplus R_R(r) = [\baddr,\eaddr]$
-    \item $\forall r \in S_R \ldotp \pwfree(r) \nsubsim
-      \stdreg{R_R(r),\gc}{\pur} \text{ and } |R_R(r)|=1$
+    \item $\forall r \in S_R \ldotp \pwfree(r).H \nsubeq
+      \stdreg{R_R(r),\gc}{\pur}.H \text{ and } |R_R(r)|=1$
     \item $\biguplus R_W(r) = [\baddr,\eaddr]$
-    \item $\forall r \in S_W \ldotp \pwfree(r) \nsupsim
-      \stdreg{R_W(r),\gc}{\pur} \text{ and } |R_R(r)|=1 \text{ and } \pwfree(r)
+    \item $\forall r \in S_W \ldotp \pwfree(r).H \nsupeq
+      \stdreg{R_W(r),\gc}{\pur}.H \text{ and } |R_R(r)|=1 \text{ and } \pwfree(r)
       \text{ is address-stratified}$
@@
   \]
-  Assuming for contradiction $\exists r \in S \ldotp R_R(r) \neq R_W(r)$, we have $\pwfree(r) \nsubsim \stdreg{*,\gc}{\pur}$ and $\pwfree(r) \nsupsim \stdreg{-,\gc}{\pur}$ for two distinct singleton sets $*$ and $-$.
+  Assuming for contradiction $\exists r \in S \ldotp R_R(r) \neq R_W(r)$, we have $\pwfree(r).H \nsubeq \stdreg{*,\gc}{\pur}.H$ and $\pwfree(r).H \nsupeq \stdreg{-,\gc}{\pur}.H$ for two distinct singleton sets $*$ and $-$.
   Due to the memory satisfaction assumption $\pwfree(r)$ is non-empty and it is trivial to show that so is the standard regions.
@@
   \label{lem:nsub-and-nsup-std}
-  If $\iota \nsubsim \stdreg{A,\gc}{v}$ and $\iota \nsupsim \stdreg{A,\gc}{v}$, then $\iota \nsim {A,\gc}{v}$.
+  If $\iota.H \nsubeq \stdreg{A,\gc}{v}.H$ and $\iota.H \nsupeq \stdreg{A,\gc}{v}.H$, then $\iota.H \nequal  \stdreg{A,\gc}{v}.H$
 \end{lemma}

### 2018-06-07 fully-abstract-comp/technical-report.tex
Now require exec capabilities to be normal in V-relation. Updated execCond to no longer depend on linearity.

@@
         %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
-        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{\lin,W} \wedge\\
-                                  \npair{[\baddr,\eaddr]} \in \xReadCond{W}\endarray
+        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
+                                  \npair{[\baddr,\eaddr]} \in \xReadCond{W}
+                                  \lin = \normal \wedge \\ \endarray
       \end{array}
@@
 \begin{align*}
-  \execCond{\lin,W} &=
+  \execCond{W} &=
   \left\{ \npair{A} \middle|
@@
       \forall n' < n, W' \future \purePart{W}, \aaddr \in [\baddr',\eaddr'] \subseteq A \ldotp\\
-      \quad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
-    \end{array}
-    \right\}\\
-  &\cup\left\{ \npair{A} \middle|
-    \begin{array}{l}
-      \lin = \linear \wedge\\
-      \exists W_a : A \fun \World \wedge W = \oplus_{a\in[\baddr,\eaddr]} W_a(a)\wedge\\
-      \forall n' < n, \aaddr\in[\baddr',\eaddr'] \subseteq A, W' \future \oplus_{a\in[\baddr',\eaddr']} W_a(a) \ldotp\\
-      \quad \npair[n']{\stpair[.]{((\rx,\lin),\baddr',\eaddr',\aaddr)}{((\rx,\lin),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+      \quad \npair[n']{\stpair[.]{((\rx,\normal),\baddr',\eaddr',\aaddr)}{((\rx,\normal),\baddr',\eaddr',\aaddr)}} \in \lre(W')
     \end{array}
@@
   \item If $\npair{A} \in \writeCond{\normal,W}$, then $\npair{A} \in \writeCond{}(\normal,\purePart{W})$
-  \item If $\npair{A} \in \execCond{\normal,W}$, then $\npair{A} \in \execCond{}(\normal,\purePart{W})$
+  \item If $\npair{A} \in \execCond{W}$, then $\npair{A} \in \execCond{}(\purePart{W})$
   \item If $\npair{(w_1,w_2)} \in \lrv(H_\sigma,W)$ and ($\nonLinear{w_1}$ or
@@
     $\npair{A'} \in \xReadCond{W}$.
-  \item If $\npair{A} \in \execCond{\lin,W}$  and $\emptyset \neq A' \subseteq A$, then
-    $\npair{A'} \in \execCond{\lin,W}$.
+  \item If $\npair{A} \in \execCond{W}$  and $\emptyset \neq A' \subseteq A$, then
+    $\npair{A'} \in \execCond{W}$.
   \end{itemize}
@@
   \item If $\npair{A} \in \writeCond{\lin,W}$, then there exists $W_1, W_2$ such that $W = W_1\oplus W_2$ such that $\npair{[\baddr_1,\eaddr_1]} \in \writeCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \writeCond{\lin,W_2}$.
-  \item If $\npair{A} \in \execCond{\lin,W}$, then there exists $W_1, W_2$ such that $W = W_1\oplus W_2$ such that $\npair{[\baddr_1,\eaddr_1]} \in \execCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \execCond{\lin,W_2}$.
+  \item If $\npair{A} \in \execCond{W}$, then there exists $W_1, W_2$ such that $W = W_1\oplus W_2$ such that $\npair{[\baddr_1,\eaddr_1]} \in \execCond{W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \execCond{W_2}$.
   \end{itemize}
@@
       by assumption and Lemma~\ref{lem:conds-shrinkable}.
-    \item $\npair{[b,e]}\in\execCond{\normal,W}$:
+    \item $\npair{[b,e]}\in\execCond{W}$:
       take $n' < n$, $W' \future \purePart{W}$, $\aaddr' \in [\baddr',\eaddr'] \subseteq [\baddr,\eaddr]$ and $w' = ((\rx,\normal),\baddr',\eaddr',\aaddr')$.

### 2018-06-07 fully-abstract-comp/ftlr-proof-cases.tex
return pointer data case for FTLR call case.

@@
     \item $\Phi_S''.\mem = \Phi_S.\mem$
-    \item $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{\aaddr_\stk,\eaddr_\stk}\update{\aaddr_\stk}{42}$
+    \item $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$
     \item $\Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk$
@@
 \[
-  W_{R,1}' = W_{R,1}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto \stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square}]
+  W_{R,1}' = W_{R,1}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)]
 \] 
@@
   \[
-    \npair[n'-1]{(\sealed{\sigma',\retptrc(\aaddr_\stk,\eaddr_\stk)},\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk)})} \in \lrv(W_{R,\rretd}')
+    \npair[n'-1]{(\sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)},\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)})} \in \lrv(W_{R,\rretd}')
   \]
   where $W_{R,\rretd}'$ is the part of $W_R'$ with ownership over $r_\var{priv\_stk}$.
-  From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get a
-  set of regions $R \subseteq \addressable{\normal, \pwheap}$ 
-and
-  $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W}$,
+  From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}$.
+  Further by $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one
+  $r \in S$ there is a $\mscode' \subseteq \mscode$\footnote{Follows from
+    $H^\var{code}$ requiring $\dom(\mscode') \subseteq \ta$ and $\dom(\mscode) =
+    \ta$ and finally that $\mscode$ is part of $\ms_S$.} such that
+  \begin{equation}
+    \npair[n']{(\mscode',\mscode')} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\_)\label{eq:ftlr:rretd:Hcode}
+  \end{equation}
+ where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
+  That is $\mscode'$ contains the call we are considering.
+  \lau{TODO: Make the above more precise and possibly put into a lemma.}
+  Use the $r$ in question as the witness. We now have to show:
+  \begin{enumerate}[label=\alph*)]
+  \item for $n'' < n'-1$ we have
+    \[\npair[n'']{\left(\array{l}\retptrd(\aaddr_\stk,\eaddr_\stk),\\((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)\endarray\right)}
+      \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode'
+      \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_{R,\rretd}')
+    \]
+    First $\dom(\mscode') \subseteq \ta$ follows from \ref{eq:ftlr:rretd:Hcode} and the assumption $[\baddr,\eaddr] \ subseteq \ta$.
+    Now pick $r_\var{priv\_stk}$ as the witness.
+    We immediately get $\dom(\ms_{\stk\_\priv,S}) = \dom(\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}) = [\aaddr_\stk,\eaddr_\stk]$.
+    Next, $\decInstr(\mscode'(\aaddr,\aaddr+\calllen-1)) = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ follows from $\callCond{\Phi_S,r_1,r_2,\aaddr}$.
+    Finally, $\ms_\code(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr}$ with $\sigma' = \sigma_\baddr + \offpc \in \sigrets'$ which follows from $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$ and the fact that the call is there.
+  \item $\isLinear{\sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)}}$ iff
+    $\isLinear{\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk)}}$\\
+    Trivial as both are linear.    
+  \item
+    \[
+      \begin{array}{l}
+        \forall W'' \future \purePart{W_{R,\rretd}'}, W_o, n'' < n'-1,\\
+        \quad\npair[n'']{(\vsc_S',\vsc_T')}
+        \in  H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode'
+        \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o) \ldotp \\
+        \qquad\npair[n']{\retptrd(\aaddr_\stk,\eaddr_\stk),\vsc_S',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1),\vsc_T'} \in
+        \lrexj(W'\oplus W_o))
+      \end{array}
+    \]
+    Trivial as both configuration fails.
+  \end{enumerate}
 
@@
 \end{description}
+
+Case: \ref{case:ftlr:scall:mem}, \lau{TODO}
 

### 2018-06-11 fully-abstract-comp/technical-report.tex
return pointer data case for FTLR call case.
Fix typo.

@@
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
-      \forall a \in \dom(\code), n' < n \ldotp\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})
+      \forall a \in \dom(\code), n' < n \ldotp\\
+      \quad\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})
     \end{array}
@@
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \; \hat{W} = \\
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  (\ta,\stkb) \; \sigma \; \hat{W}= \\
   \begin{array}[t]{l}
@@
         \dom(\code) \subseteq \ta \tand\\
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms,\ms),(\ta,\stkb)]{\spa,\square}, \aaddr'+\calllen) \tand \\
-        \quad \dom(\ms) = [\baddr,\eaddr] \tand\\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
         \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
@@
   \begin{itemize}
-  \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) or ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
+  \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) and ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
   \end{itemize}

### 2018-06-11 fully-abstract-comp/ftlr-proof-cases.tex
Big chunk of work on component-related lemmas
* Refactor heap satisfaction judgement to take the heap regions and world
  separately.
* Remove "World satisfies component" judgement.
* Instead, make the component FTLR produce a suitable world, that the code and
  data memory satisfy.
* Update the rest of the tech report to follow suit.

@@
     \lrv(W')$.
+  \item If $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap(\pwheap[W_1])(W)$, then $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in
+    \lrheap(\pwheap[W_1])(W')$.
 \end{itemize}
 \end{lemma}
+\begin{proof}
+  Follows from the definitions.
+  Note: the proof for $\lrheap$ relies on Lemma~\ref{lem:oplus-future-distr}.
+\end{proof}
 
@@
   \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
-  \item $\memSatHeap{\ms_S',\ms_\var{T,heap}}{W_\var{heap}}$.
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(W_M)(W_\var{heap})$.
   \end{itemize}
 
-  From $\memSatHeap{\ms_S,\ms_\var{T,heap}}{W_\var{heap}}$, we get an
+  From $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(W_M)(W_\var{heap})$, we get an
   $R_\ms : \dom(\activeReg{\pwheap[W_\var{heap}]}) \fun \MemSeg \times \MemSeg$,
@@
   $W_\var{heap} = \oplus_{r \in \dom(\activeReg{\pwheap[W_\var{heap}]})} R_W(r)$,
-  $\forall r \in \activeReg{\pwheap[W_\var{heap}]}$, we have that
+  $\forall r \in \activeReg{\pwheap[W_M]}$, we have that
   $\npair{R_\ms(r)} \in  \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
+  We also get an $R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal}$ such that  $\biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)$.
 
@@
   \activeReg{\pwheap[W_\var{heap}]}$, so $\npair{R_\ms(r)} \in
-  \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
-  Because $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$ and $W\oplus W_M = W \oplus (W_\var{heap} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that
-  also $\pwheap[W_\var{heap}](r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$. This
+  \pwheap[W_M](r).H \; \xi^{-1}(R_W(r))$.
+  Because $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$ and $W\oplus W_M$ is defined, it follows that
+  also $\pwheap[W_M](r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$. This
   means that also $\npair{R_\ms(r)} \in
@@
   Additionally, if $\npair{(b,e)}\in\writeCond{l,W}$, then
-  we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_\var{heap}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
+  we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_M]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
   for all $r \in S'$, $\pwheap(r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$ and
@@
   Because $W \oplus W_M$ is defined, it follows that also 
-  $\pwheap[W_\var{heap}](r').H \nsupeq \stdreg{R'(r'),\gc}{\pur}.H$ and
-  $\pwheap[W_\var{heap}](r') \text{ is address-stratified}$.
+  $\pwheap[W_M](r').H \nsupeq \stdreg{R'(r'),\gc}{\pur}.H$ and
+  $\pwheap[W_M](r') \text{ is address-stratified}$.
 
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_\var{heap}](r)$ is address-stratified and $\pwheap[W_\var{heap}](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
+  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_M](r)$ is address-stratified and $\pwheap[W_M](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
 
-  From this, it follows that $\memSatHeap{\ms_S\update{a}{0},\ms_\var{T,heap}\update{a}{0}}{W_r' \oplus W_\var{heap}'}$ and finally $\memSat{(\ms_S\update{a}{0},\stk,\ms_\stk,\ms_T\update{a}{0})}{W_M'}$.
+  From this, it follows that $\npair{(\overline{\sigma},\ms_S\update{a}{0},\ms_\var{T,heap}\update{a}{0})} \in \lrheap(W_M)(W_r' \oplus W_\var{heap}')$ and finally $\npair{(\overline{\sigma},\ms_S\update{a}{0},\stk,\ms_\stk,\ms_T\update{a}{0})} \in \lrheap(W_M')(W_r' \oplus W_{\var{heap}}')$.
 \end{proof}
@@
   \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
-  \item $\memSatHeap{\ms_S',\ms_\var{T,heap}}{W_\var{heap}}$.
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})}\in\lrheap(\pwheap[W_M])(W_\var{heap})$.
   \end{itemize}
@@
   \item \begin{itemize}
-    \item $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$
     \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$

### 2018-06-11 fully-abstract-comp/technical-report.tex
Big chunk of work on component-related lemmas
* Refactor heap satisfaction judgement to take the heap regions and world
  separately.
* Remove "World satisfies component" judgement.
* Instead, make the component FTLR produce a suitable world, that the code and
  data memory satisfy.
* Update the rest of the tech report to follow suit.

@@
   \inference{
-    \dom(\mscode) \mathrel{\#} \dom(\msdata)\\
+    \dom(\mscode) \mathrel{\#} \dom(\msdata)&
+    \dom(\mscode) = [\baddr,\eaddr]\\
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
@@
     \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
-    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta)
+    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)
   }{
@@
   Follows easily from the definitions.
+\end{proof}
+
+\begin{lemma}[$\oplus$ and future worlds]
+  \label{lem:oplus-future-distr}
+  If $W' \future W_1 \oplus W_2$, then there exist $W_1',W_2'$ such that $W' = W_1' \oplus W_2'$ and $W_1' \future W_1$ and $W_2' \future W_2$.
+\end{lemma}
+\begin{proof}
+  We define $W_1'$ and $W_2'$ to have the same regions as $W'$ with a possibly different visibility.
+  For regions that are present in $W_1$ and $W_2$, we give them the same visibility in $W_1'$ and $W_2'$ respectively.
+  For regions that are new in $W'$, we make them $\pure$ or $\spatial$ in both $W_1'$ and $W_2'$ if they are in $W'$ and we make them $\spatialo$ in $W_1'$ but $\spatial$ in $W_2'$ if they are spatial in $W'$.
+  It is then easy to check that the required equations hold.
 \end{proof}
@@
   \label{lem:oplus-future}
-  For all $W$, we have that:
+  For all $W_1,W_2$, we have that:
   $W_1 \oplus W_2 \future W_1$
@@
 
+\begin{lemma}[Partial authority is better than nothing]
+  \label{lem:auth-partial-betterthannothing}
+  If $W = W_1 \oplus W_2$, then $W_1 \future \purePart{W}$.
+\end{lemma}
+\begin{proof}
+  Follows easily from the definitions.
+\end{proof}
+
 \subsection{Memory satisfaction}
@@
       W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
-      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\ldotp \\
+      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}\ldotp \\
       \quad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
@@
       \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
-      \quad \memSatHeap{\ms_S',\ms_\var{T,heap}}{W_{\var{heap}}}
+      \quad \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
     \end{array}
@@
 \[
-  \memSatHeap{ms,\ms_T}{W} \text{ iff } 
+  \lrheap(\pwheap)(W') = 
   \left\{
+    \npair{(\overline{\sigma},ms,\ms_T)} \middle|
     \begin{array}{l}
@@
       \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
-      \qquad W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
-      \qquad \forall r \in \activeReg{\pwheap} \ldotp \\
+      \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{\pwheap}) \ldotp \\
       \qquad \quad \npair{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
-      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \text{ is defined } \wedge\\
+      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
       \quad \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
@@
 \begin{lemma}[Combined independent heap memory satisfies disjoint world]
-  \label{sec:combined-memory-disjoint-world}
-  If $\memSatHeap{\overline{\sigma_1},ms_{S,1},\ms_{T,1}}{W_1}$ and $\memSatHeap{\overline{\sigma_2},ms_{S,2},\ms_{T,2}}{W_2}$, then $\memSatHeap{\overline{\sigma_1}\uplus\overline{\sigma_2},ms_{S,1}\uplus \ms_{S,2},\ms_{T,1}\uplus \ms_{T,2}}{W_1 \uplus W_2}$.
+  \label{lem:combined-memory-disjoint-world}
+  If $\npair{(\overline{\sigma_1},ms_{S,1},\ms_{T,1})}\in\lrheap(\pwheap[W_1])(W)$ and $\npair{(\overline{\sigma_2},ms_{S,2},\ms_{T,2})} \in \lrheap(\pwheap[W_2])(W)$, then $\npair{(\overline{\sigma_1}\uplus\overline{\sigma_2},ms_{S,1}\uplus \ms_{S,2},\ms_{T,1}\uplus \ms_{T,2})} \in \lrheap(\pwheap[W_1 \uplus W_2])(W)$.
 \end{lemma}
 \begin{proof}
-  Unfolding the definitions, it's easy to construct the memory and seal partitions $R_{\ms,3}$ and $R_{\var{seal},3}$ from the corresponding partitions of the separate memories.
-  Constructing $R_{W,3}$ from $R_{W,1}$ and $R_{w,2}$ is a bit harder, but we can use Lemmas~\ref{lem:oplus-distr-uplus},~\ref{lem:purePart-duplicable},~\ref{lem:uplus-assoc-comm} and~\ref{lem:oplus-assoc-comm} to manipulate the world equalities as follows:
-  \begin{align*}
-    W_1 \uplus W_2 &= (\purePart{W_1} \oplus W_1) \uplus (W_2 \oplus \purePart{W_2})\\
-                   &= (\purePart{W_1} \uplus W_2) \oplus (W_1 \uplus \purePart{W_2})\\
-                   &= (\purePart{W_1} \uplus \bigoplus_r R_{W,2}(r)) \oplus ((\bigoplus_r R_{W,1}(r)) \uplus \purePart{W_2})\\
-                   &= \bigoplus_r (\purePart{W_1} \uplus R_{W,2}(r)) \oplus \bigoplus_r (R_{W,1}(r) \uplus \purePart{W_2})
-  \end{align*}
-  We can then define
-  \begin{equation*}
-    R_{W,3}(r) =
-    \begin{cases}
-      \purePart{W_1} \uplus R_{W,2}(r) &\text{when } r \in \dom(R_{W,2})\\
-      \purePart{W_2} \uplus R_{W,1}(r) &\text{when } r \in \dom(R_{W,1})
-    \end{cases}
-  \end{equation*}
-  Lemma~\ref{lem:uplus-future} and the monotonicity of the regions' memory predicates then yield the required results.
+  Unfolding the definitions, it's easy to construct the memory and seal partitions $R_{\ms,3}$ and $R_{\var{seal},3}$ and $R_{W,3}$ from the corresponding partitions of the separate memories, seals and worlds.
 \end{proof}
@@
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-    \dom(\code) \subseteq \ta \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset) \wedge \\
+    (\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
       \forall a \in \dom(\code), n' < n \ldotp\\
-      \quad\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})
+      \quad\npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
     \end{array}
@@
     \end{array}
-      \right\} \uplus\\
+    \right\} \\
+  \end{array}\\
+    \text{ if } \sigma \in \sigrets\\
+\end{multline*}
+and
+\begin{multline*}
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  (\ta,\stkb) \; \sigma \; \hat{W}= \\
   \left\{
@@
     \end{array}
-      \right\}
-  \end{array}
+  \right\}\\
+  \text{ if } \sigma \in \sigcloss
 \end{multline*}
 
-\subsection{World specifies component}
-
-\begin{definition}[World specifies component]
-  We define that a world $W$ specifies the presence of a component with global constants $\gc$
-  \begin{equation*}
-    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
-  \end{equation*}
-  (written as $W \vdash_\gc \var{comp}_0$)
-  iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that for all $i$,
-  \begin{itemize}
-  \item 
-    \begin{equation*}
-      \vdash (\code_i,\msdata[i],\overline{c_{\mathrm{import},i}},\overline{c_{\mathrm{export},i}},\sigrets_i,\sigcloss_i,A_{\linear,i})
-    \end{equation*}
-  \item $\mscode = \biguplus_i \code_i$
-  \item $\msdata = \biguplus_i \msdata[i]$
-  \item $\overline{c_{\mathrm{export}}} = \biguplus_i \overline{c_{\mathrm{export},i}}$
-  \item $\overline{c_{\mathrm{import}}} = \biguplus_i \overline{c_{\mathrm{import},i}} \setminus \overline{c_{\mathrm{export}}}$
-  \item $\sigrets = \biguplus \sigrets_i$
-  \item $\sigcloss = \biguplus \sigcloss_i$
-  \item $A_\linear = \biguplus A_{\linear,i}$
-  \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
-  \item For all $i$, one of the following holds:
-    \begin{itemize}
-    \item $\dom(\code_i) \subseteq \ta$
-    \item $\dom(\code_i) \mathrel{\#} \ta$
-    \end{itemize}
-  \end{itemize}
-  and there exist $r_{\mathrm{addr}} : \dom(\msdata) \mapsto \dom(\pwheap)$ such that
-  \begin{itemize}
-  \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) and ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
-  \end{itemize}
-  with $\gc = (\ta,\stkb)$
-\end{definition}
-Note that for untrusted components, we require all seals to be in the corresponding code region as closure seals.
-That is because we cannot trust untrusted components to treat their seals correctly (e.g.
+% \subsection{World specifies component}
+
+% \begin{definition}[World specifies component]
+%   We define that a world $W$ specifies the presence of a component with global constants $\gc$
+%   \begin{equation*}
+%     \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
+%   \end{equation*}
+%   (written as $W \vdash_\gc \var{comp}_0$)
+%   iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that for all $i$,
+%   \begin{itemize}
+%   \item 
+%     \begin{equation*}
+%       \vdash (\code_i,\msdata[i],\overline{c_{\mathrm{import},i}},\overline{c_{\mathrm{export},i}},\sigrets_i,\sigcloss_i,A_{\linear,i})
+%     \end{equation*}
+%   \item $\mscode = \biguplus_i \code_i$
+%   \item $\msdata = \biguplus_i \msdata[i]$
+%   \item $\overline{c_{\mathrm{export}}} = \biguplus_i \overline{c_{\mathrm{export},i}}$
+%   \item $\overline{c_{\mathrm{import}}} = \biguplus_i \overline{c_{\mathrm{import},i}} \setminus \overline{c_{\mathrm{export}}}$
+%   \item $\sigrets = \biguplus \sigrets_i$
+%   \item $\sigcloss = \biguplus \sigcloss_i$
+%   \item $A_\linear = \biguplus A_{\linear,i}$
+%   \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
+%   \item For all $i$, one of the following holds:
+%     \begin{itemize}
+%     \item $\dom(\code_i) \subseteq \ta$
+%     \item $\dom(\code_i) \mathrel{\#} \ta$
+%     \end{itemize}
+%   \end{itemize}
+%   and there exist $r_{\mathrm{addr}} : \dom(\msdata) \mapsto \dom(\pwheap)$ such that
+%   \begin{itemize}
+%   \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) and ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
+%   \end{itemize}
+%   with $\gc = (\ta,\stkb)$
+% \end{definition}
+% Note that for untrusted components, we require all seals to be in the corresponding code region as closure seals.
+% That is because we cannot trust untrusted components to treat their seals correctly (e.g.
 not hand them out to other components, only sign return capabilities with return seals etc.), so instead, they must be considered as closure seals, which the component is essentially allowed to sign anything with.
 
-\begin{lemma}[World specifies linked component if it specifies each separately]
-  \label{lem:world-specifies-linked}
-  If
-  \begin{itemize}
-  \item $W \vdash_\gc \var{comp}_0$
-  \item $W \vdash_\gc \var{comp}_0'$
-  \end{itemize}
-  Then $W \vdash \var{comp}_{0} \bowtie \var{comp}_{0}'$
-\end{lemma}
-\begin{proof}
-  Follows easily from the definitions.
-\end{proof}
-
-\begin{lemma}[World specifies component pure]
-  \label{lem:world-specifies-comp-pure}
-  If $W \vdash_\gc \var{comp}_0$ then $\purePart{W} \vdash \var{comp}_{0}$.
-\end{lemma}
-\begin{proof}
-  Follows easily from the definitions.
-\end{proof}
+% \begin{lemma}[World specifies linked component if it specifies each separately]
+%   \label{lem:world-specifies-linked}
+%   If
+%   \begin{itemize}
+%   \item $W \vdash_\gc \var{comp}_0$
+%   \item $W \vdash_\gc \var{comp}_0'$
+%   \end{itemize}
+%   Then $W \vdash \var{comp}_{0} \bowtie \var{comp}_{0}'$
+% \end{lemma}
+% \begin{proof}
+%   Follows easily from the definitions.
+% \end{proof}
+
+% \begin{lemma}[World specifies component pure]
+%   \label{lem:world-specifies-comp-pure}
+%   If $W \vdash_\gc \var{comp}_0$ then $\purePart{W} \vdash \var{comp}_{0}$.
+% \end{lemma}
+% \begin{proof}
+%   Follows easily from the definitions.
+% \end{proof}
 
@@
   \item \begin{itemize}
-    \item $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss,\_)$
     \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
@@
   \lrcomp(W) &=
-  \left\{\begin{multlined}
-     \npair{\var{comp},\var{comp}} \;\mid \;\\
-      \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
-      W \vdash \var{comp} \tand\\
-      \text{For all } W' \future W \ldotp \text{If } \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
-      \text{then } \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
-  \end{multlined}
+  \left\{\begin{aligned}
+     &\npair{\var{comp},\var{comp}} \;\mid \;\\
+      &\qquad\var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+      &\qquad\text{For all } W' \future W \ldotp \\
+      &\qquad\quad\text{If } \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W') \tand\\
+      &\qquad\quad\overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+  \end{aligned}
     \right\}\\
@@
   Therefore, we can restrict ourselves to the case where both pairs of components are of the form $\var{comp}_0$, i.e. no ``main'' capabilities.
-
-  By Lemma~\ref{lem:world-specifies-linked}, we have that $W \vdash \var{comp}_1\bowtie \var{comp}_2$.
 
@@
   \item $A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}$
-  \end{itemize}
-
-  Now take $W' \future W$ and assume that $\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
-  Then it remains to show that $\overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$.
+  \item $\dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3])$
+  \item $\sigrets[3] \mathrel{\#} \sigcloss[3]$
+  \end{itemize}
+
+  Now take $W' \future (W_1\uplus W_2)$ and assume that $\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
+  Then it remains to show that
+  \begin{equation*}
+    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3], \mscode[3]\uplus\msdata[3])} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+  \end{equation*}
+  and
+  \begin{equation*}
+    \overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})
+  \end{equation*}
   We do this by complete induction on $n$, so that we can assume that $\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
 
-  For each $$\npair{(c_{\mathrm{export}},c_{\mathrm{export}})} \in \overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}}$$, we have that $\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}$ is either in $\overline{\npair{(c_{\mathrm{export},1},c_{\mathrm{export},1})}}$ or $\overline{\npair{(c_{\mathrm{export},2},c_{\mathrm{export},2})}}$.
-  By unfolding $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W)$, it suffices to prove that $$\overline{\npair[n']{(c_{\mathrm{import},1},c_{\mathrm{import},1})}} \in \lrv(\purePart{W'})$$ and $$\overline{\npair[n']{(c_{\mathrm{import},2},c_{\mathrm{import},2})}} \in \lrv(\purePart{W'})$$
+  First, we prove that
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import},1},c_{\mathrm{import},1})}} \in \lrv(\purePart{W'})
+  \end{equation*}
+  and
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import},2},c_{\mathrm{import},2})}} \in \lrv(\purePart{W'})
+  \end{equation*}
   for all $n' < n$.
-
   Simple set algebra tells us that 
-  $$\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}} \subseteq \overline{c_{\mathrm{export},3}} \cup \overline{c_{\mathrm{import},3}}$$
-  and for all $n' < n$, we know that $$\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$$
+  \begin{equation*}
+    \overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}} \subseteq \overline{c_{\mathrm{export},3}} \cup \overline{c_{\mathrm{import},3}}
+  \end{equation*}
+  and for all $n' < n$, we know that
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})
+  \end{equation*}
   and
-  $$\overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})\text{.}$$
+  \begin{equation*}
+    \overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})\text{.}
+  \end{equation*}
+
+  Then, it follows from our assumptions ($\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$) that
+  \begin{align*}
+    \npair{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1], \mscode[1]\uplus\msdata[1])} &\in \lrheap(\pwheap[W_1])(W')\\
+    \npair{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2], \mscode[2]\uplus\msdata[2])} &\in \lrheap(\pwheap[W_2])(W')
+  \end{align*}
+  It follows by Lemma~\ref{lem:combined-memory-disjoint-world} that 
+  \begin{equation*}
+    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3], \mscode[3]\uplus\msdata[3])} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+  \end{equation*}
+
+  Finally, for each
+  \begin{equation*}
+    \npair{(c_{\mathrm{export}},c_{\mathrm{export}})} \in \overline{\npair{(c_{\mathrm{export},3},c_{\mathrm{export},3})}}
+  \end{equation*}
+  we have that $\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}$ is either in $\overline{\npair{(c_{\mathrm{export},1},c_{\mathrm{export},1})}}$ or $\overline{\npair{(c_{\mathrm{export},2},c_{\mathrm{export},2})}}$.
+  By unfolding $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$ and using Lemma~\ref{lem:uplus-future}, this follows from the results above.
 \end{proof}
@@
 
+\begin{lemma}[Code region sealing invariant for untrusted components implies untrusted safety]
+  \label{lem:codereg-sealing-safety}
+  If
+  \begin{itemize}
+  \item $\emptyset\neq\dom(\mscode) \mathrel{\#} \ta$
+  \item $\sigma \in \sigcloss$
+  \end{itemize}
+  then $\codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \sigma \nequal \lrv$ for all $n$
+\end{lemma}
+\begin{proof}
+  The result follows easily by definition of $\codereg{}$ and $H^\mathrm{code,\square}_\sigma$.
+\end{proof}
+
 \begin{lemma}[non-linear words are pure]
@@
     By definition of $\lrvg{\trusted}(W)$ and Lemma~\ref{lem:conds-shrinkable} with the assumption that $\npair{\dom(\mscode)}\in\xReadCond{W}$.
+  \end{itemize}
+\end{proof}
+
+\begin{lemma}[FTLR for component code-values]
+  \label{lem:ftlr-comp-code}
+  If
+  \begin{equation*}
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} w
+  \end{equation*}
+  and
+  \begin{itemize}
+  \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item ($\dom(\mscode) \mathrel{\#} \ta$ and $\trust = \untrusted$) or ($\dom(\mscode) \subseteq \ta$ and $\trust = \trusted$)
+  \end{itemize}
+
+  then $\npair{(w,w)} \in \lrvg{\trust}(W)$
+\end{lemma}
+\begin{proof}
+  By induction on
+  \begin{equation*}
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} w
+  \end{equation*}
+  There are two cases to consider:
+  \begin{itemize}
+  \item
+    $\mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ and $[\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)$:
+
+    We distinguish two cases:
+    \begin{itemize}
+    \item $\trust = \trusted$:
+
+      By definition of $\lrvg{\trusted}$, it suffices to prove that
+      $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$, $\dom(\code)\subseteq \ta$ and $[\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)$, all of which follow by assumption.
+
+    \item $\trust = \untrusted$:
+
+      In this case, we know that $\sigrets=\emptyset$.
+
+      By definition of $\lrv$, it suffices to prove that
+      \begin{equation*}
+        \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal \lrv
+      \end{equation*}
+      We take $r = r_{\var{code}}$ and the result then follows from Lemma~\ref{lem:codereg-sealing-safety}.
+    \end{itemize}
+
+  \item $\mscode(a) \in \ints$ and
+    \begin{equation*}
+      \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
+    \end{equation*}
+    Using the former, the result follows easily by definition.
   \end{itemize}
@@
     \end{itemize}
-  \item $W \vdash \var{comp}$
-  \end{itemize}
-  Then $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$.
+  \end{itemize}
+  Then there exists a $W$ such that
+  \begin{itemize}
+  \item $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$
+  \item $\dom(\pwheap)$ can be chosen to not include any finite set of region names 
+  \item $\dom(\pwfree) =
+    \dom(\pwpriv)=\emptyset$
+  \end{itemize}
 \end{lemma}
@@
   Hence, we can restrict our attention to components of the form $\var{comp}_0$.
+
   Take
@@
   \end{equation*}
-
-  We know that $W \vdash \var{comp}$ by assumption, so take $W' \future W$ and
+  We then know from $\vdash \var{comp}$ that 
+  \begin{itemize}
+  \item $\dom(\mscode) \mathrel{\#} \dom(\msdata)$
+  \item $\dom(\mscode) = [\baddr,\eaddr]$
+  \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$
+  \item $\exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}$
+  \item $\dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear$
+  \item $A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a)$
+  \item
+    \begin{equation*}
+      \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)
+    \end{equation*}
+  \item
+    \begin{equation*}
+      \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}
+    \end{equation*}
+  \item $\overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}}$
+  \item $(\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)$
+  \end{itemize}
+
+  Take $W$ such that
+  \begin{itemize}
+  \item $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$
+  \item $\pwheap(r_\code) =  \codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}$
+  \item there exist $r_{\mathrm{addr}} : \dom(\msdata) \mapsto \dom(\pwheap)$ such that
+    \begin{itemize}
+    \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) and ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
+    \end{itemize}
+  \end{itemize}
+  with $r_{\mathrm{code}}$ and $r_a$ chosen according to the given restriction.
+
+  It now remains to prove that $\npair{(\var{comp},\var{comp})} \in \lrcomp{W}$.
+  Take $W' \future W$ and assume that
   \begin{equation*}
@@
   \end{equation*}
-  then it remains to prove that
-  \begin{equation*}
-    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
-  \end{equation*}
-
-  We prove by general induction on $n$ that this is the case for all $\overline{c_{\mathrm{export}}}$, i.e. assume that for all $n' < n$,
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
-  \end{equation*}
-
-  Because $W \vdash \var{comp}$, for every $c_{\mathrm{export}} \in \overline{c_{\mathrm{export}}}$, there must exist an $i$ such that
+  We need to show that
   \begin{itemize}
@@
     \begin{equation*}
-      \var{comp}_i = (\code_i,\msdata[i],\overline{c_{\mathrm{import},i}},\overline{c_{\mathrm{export},i}},\sigrets_i,\sigcloss_i,A_{\linear,i})
+      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
     \end{equation*}
-  \item $\vdash \var{comp}_i$
-  \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
-  \item One of the following holds:
-    \begin{itemize}
-    \item $\dom(\code_i) \subseteq \ta$
-    \item $\dom(\code_i) \mathrel{\#} \ta$
-    \end{itemize}
-  \item $c_{\mathrm{export}} \in \overline{c_{\mathrm{export},i}}$
-  \end{itemize}
-
-  It follows easily that for any $n$,
-  \begin{itemize}
-  \item $\npair{A_{\linear,i}} \in \readCond{\linear,W}$,
-  \item $\npair{A_{\linear,i}} \in \writeCond{\linear,W}$,
-  \item $\npair{A_{\mathrm{non-linear},i}} \in \readCond{\normal,W}$ and
-  \item $\npair{A_{\mathrm{non-linear},i}} \in \writeCond{\normal,W}$
-  \end{itemize}
-  with $A_{\mathrm{non-linear},i} = \dom(\msdata[i]) \setminus A_{\linear,i}$
-
-  We have that the component is well-formed: $\vdash \var{comp}_i$, which implies that
-  \begin{equation*}
-    \dom(\mscode[i]), \emptyset, A_{\mathrm{non-linear},i}, \overline{c_{\mathrm{import},i}}, \sigrets_i,\sigcloss_i \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}
-  \end{equation*}
-  and $\overline{c_{\mathrm{export},i}} \mathrel{\#} \overline{c_{\mathrm{import},i}}$.
-
-  Using Lemma~\ref{lem:ftlr-comp-value} and the above results, it then remains to prove that
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import},i},c_{\mathrm{import},i})}} \in \lrv(\purePart{W}) \text{ for all } n' < n
-  \end{equation*}
-
-  From $W \vdash \var{comp}$, we know that $\overline{c_{\mathrm{import},i}} \subseteq \overline{c_{\mathrm{import}}} \cup \overline{c_{\mathrm{export}}}$.
-  For the former, we know by assumption that 
+   Take $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that
+    \begin{itemize}
+    \item $R_\ms(r_\addr) = \msdata|_{\{a\}} \times \msdata|_{\{a\}}$.
+    \item $R_\ms(r_\var{code}) = \mscode\times\mscode$
+    \end{itemize}
+    Take $R_W' : \dom(\activeReg{\pwheap}) \fun \World$ such that
+    \begin{itemize}
+    \item $R_W'(r_\addr) = \purePart{W}[\var{heap}.r_{\addr'}\mapsto \stdreg{\{\aaddr\},\gc}{\spatialo}]_{\addr' \in A_\var{own}(\addr)}$ 
+    \item $R_W'(r_\var{code}) = \purePart{W}$
+    \end{itemize}
+    Since then $W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W'(r)$, we can use Lemma~\ref{lem:oplus-future-distr} to construct an $R_W$ with  $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $R_W(r) \future R_W'(r)$ for all $r$.
+    Finally, take $R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal}$ to satisfy
+    \begin{itemize}
+    \item $R_\var{seal}(r_\addr) = \emptyset$
+    \item $R_\var{seal}(r_\var{code}) = \sigrets\uplus \sigcloss$.
+    \end{itemize}
+
+    We then need to prove that 
+    \begin{itemize}
+    \item $\npair{(\msdata|_{\{\addr\}},\msdata|_{\{\addr\}})} \in \stdreg{\{\aaddr\},\gc}{\lin}.H~ \xi^{-1}(R_W(r_\addr))$:
+      By definition, it suffices to prove that for all $n' < n$,
+      $\npair[n']{(\msdata(\addr),\msdata(\addr))} \in \lrv(R_W(r_\addr))$.
+
+      We can assume that $\msdata(\addr) \not\in \overline{c_{\var{import}}}$, because otherwise the result follows from the previous assumption that 
       \begin{equation*}
@@
       \end{equation*}
-  and for the latter, our induction hypothesis tells us that
+      together with \ref{lem:monotonicity} and the fact that $R_W(\addr) \future \purePart(W')$ (by Lemma~\ref{lem:auth-partial-betterthannothing}).
+
+      We know from our assumptions that
       \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
+        \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)
       \end{equation*}
+
+      By Lemma~\ref{lem:ftlr-comp-value}, it then suffices to prove that
+      \begin{itemize}
+      \item $\npair{A_{\mathrm{non-linear}}} \in \readCond{\normal,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair{A_{\mathrm{non-linear}}} \in \writeCond{\normal,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair{A_{\mathrm{own}}(\addr)} \in \readCond{\linear,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair{A_{\mathrm{own}}(\addr)} \in \writeCond{\linear,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\pwheap[R_W(\addr)](r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$: follows from the choice of $W$ and $R_W'(\addr)$ and the fact that $R_W(a) \future R_W'(a)$
+      \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$: by assumption.
+      \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(\purePart{R_W(\addr)})$ for all $n' < n$: by the fact that $\purePart{R_W(\addr)} = \purePart{W'}$ and the above assumption that .
+        \begin{equation*}
+          \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$} 
+        \end{equation*}
+      \item $w \not\in \overline{c_{\mathrm{import}}}$: by assumption.
+      \end{itemize}
+    \item $\npair{(\mscode,\mscode)} \in  \codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H~ \xi^{-1}(\purePart{W'})$:
+      We take $\trust = \trusted$ if $\dom(\code) \subseteq \ta$ and $\trust =\untrusted$ otherwise.
+      By definition, we need to prove that:
+
+      \begin{itemize}
+      \item $\dom(\code) = [\aaddr,\baddr]$: by assumption.
+      \item $(\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)$: by assumption and choice of $\trust$.
+      \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code$: by assumption.
+      \item $\forall a \in \dom(\code), n' < n \ldotp \npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{}(\xi(\xi^{-1}(\purePart{W'}))))$:
+        Note first that $\purePart{}(\xi(\xi^{-1}(\purePart{W'}))) =\purePart{W'}$.
+
+        By Lemma~\ref{lem:ftlr-comp-code}, it suffices to prove that:
+        \begin{itemize}
+        \item $\pwheap[W'](r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$: follows by choice of $W$ and the fact that $W' \future W$.
+        \item ($\dom(\mscode) \mathrel{\#} \ta$ and $\trust = \untrusted$) or ($\dom(\mscode) \subseteq \ta$ and $\trust = \trusted$): by assumption and choice of $\trust$
+        \end{itemize}
+      \end{itemize}
+    \item
+      \begin{equation*}
+        \dom(\codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma) = \sigrets\uplus \sigcloss
+      \end{equation*}
+      This follows easily from the definition.
+  \end{itemize}
+  \item $\overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})$: TODO
+
+    We know from our assumptions that
+    \begin{equation*}
+      \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}
+    \end{equation*}
+
+    By Lemma~\ref{lem:ftlr-comp-value}, it then suffices to prove that
+    \begin{itemize}
+    \item $\npair{A_{\mathrm{non-linear}}} \in \readCond{}(\normal,\purePart{W'})$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+    \item $\npair{A_{\mathrm{non-linear}}} \in \writeCond{}(\normal,\purePart{W'})$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+    \item $\pwheap[\purePart{W'}](r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$: follows from the choice of $W$ and the fact that $W' \future W$
+    \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$: by assumption.
+    \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(\purePart{}(\purePart{W'}))$ for all $n' < n$: by the fact that $\purePart{}(\purePart{W'}) = \purePart{W'}$ (Lemma~\ref{lem:purePart-idempotent}) and the above assumption that
+      \begin{equation*}
+        \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$} 
+      \end{equation*}
+    \item $w \not\in \overline{c_{\mathrm{import}}}$: by assumption.
+    \end{itemize}
+  \end{itemize}
 \end{proof}
+
 Note: the trusted case of the above lemma can be considered as a compiler correctness result.
@@
   $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
-  From this, it follows that $W \vdash \var{comp}$ and for all  $W' \future W$, if
+  From this, it follows that for all  $W' \future W$, if
   \begin{equation*}

### 2018-06-11 fully-abstract-comp/technical-report.tex
Progress on component-related lemmas.

@@
 
+\begin{lemma}[purePart respects $\oplus$]
+  \label{lem:purePart-oplus}
+ For all $W_1,W_2$, we have that $\purePart{W_1\oplus W_2} = \purePart{W_1} \oplus \purePart{W_2}$.
+\end{lemma}
+\begin{proof}
+  Follows from the definition of $\purePart{}$ and $\oplus$.
+\end{proof}
+
 \begin{lemma}[purePart is monotone]
@@
 \begin{align*}
-  \lrexjg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
     \begin{array}{l}
@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\}) (W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\untrusted}(\{\rdata\}) (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
@@
       \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
-      \reg_S(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr) \tand (\trust = \trusted \text{ iff } \aaddr \in \ta) \tand \\
-      \npair{( \reg_S(\pcreg),\reg_T(\pcreg) )} \in \lrvg{\trust}(W_\pcreg) \tand \\
+      \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
       \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
-      \npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)
+      \npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\})(W_R)
     \end{array}
@@
   \begin{itemize}
-  \item $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$
+  \item $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$
+  \item $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$
   \item $\var{comp} \rightsquigarrow \Phi$,
-  \item TODO: no stack regions yet?
   \end{itemize}
@@
   \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W')$
+  \item $\npair{(\Phi,\Phi)}\in\lrec(W')$
   \end{itemize}
@@
   \item $\overline{c_{\mathrm{import}}} = \emptyset$
-  \end{itemize}
-
-  From $\npair{\stpair{\var{comp}}{\var{comp}}} \in \lrcomp(W)$, we know that
+  \item $\Phi = (\mem, \reg\src{, \emptyset, \ms_\stk})$
+  \end{itemize}
+
+  From $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$, we know that
   $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and 
@@
   \begin{equation*}
+    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
+  \end{equation*}
+  and
+  \begin{equation*}
     \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
   \end{equation*}
-  Since $\overline{c_{\mathrm{import}}} = \emptyset$, the former holds vacuously and the latter follows for every $W' \future W$.
-
-  Now take $\overline{r_a} \not\in\dom(\pwfree)$ and 
+  Since $\overline{c_{\mathrm{import}}} = \emptyset$, the former holds vacuously and the latter two results follow for every $W' \future W$.
+
+  Now take $\overline{r_a}$ for $a \in [\baddr_\stk,\eaddr_\stk]$ arbitrary and 
   \begin{align*}
-    W' &= W\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\pur}}}_{a \in [\baddr_\stk,\eaddr_\stk]}\\
+    \pwheap[W_\stk] &= \emptyset\\
+    \pwpriv[W_\stk] &= \emptyset\\
+    \dom(\pwfree[W_\stk]) &= \{r_a \mid a \in [\baddr_\stk,\eaddr_\stk]\}\\
+    \pwfree[W_\stk](r_a) &= \stdreg{\{a\},\gc}{\spatialo}\\
+    W' &= W \uplus W_\stk\\
     W_\pcreg &= \purePart{W'}\\
-    W_R &= \purePart{W}\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\spatialo}}}_{a \in [\baddr_\stk,\eaddr_\stk]}\\
-    W_M &= W\overline{\update{\mathrm{free}(r_a)}{\stdreg{\{a\},\gc}{\spatial}}}_{a \in [\baddr_\stk,\eaddr_\stk]}
+    W_R &= \purePart{W} \uplus W_\stk\\
+    W_M &= W \uplus \purePart{W_\stk}
   \end{align*}
-  Then $W' = W_M \oplus W_R \oplus W_\pcreg$ and by definition of $\lrec$, it suffices to prove that 
-  \begin{itemize}
-  \item $\reg_S(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr)$:
-    TODO
-  \item $(\trust = \trusted \text{ iff } \aaddr \in \ta)$:
-    TODO
-  \item $\npair{( \reg_S(\pcreg),\reg_T(\pcreg) )} \in \lrvg{\trust}(W_\pcreg)$:
+  Take $\trust = \trusted$ iff $a \in \ta$. 
+  Then $W' = W_M \oplus W_R \oplus W_\pcreg$ (by Lemmas~\ref{lem:purePart-duplicable},~\ref{lem:oplus-distr-uplus},~\ref{lem:oplus-assoc-comm} and~\ref{lem:uplus-assoc-comm}), $W_M \future W$ (by Lemma~\ref{lem:uplus-future}) and $W_\pcreg \future \purePart{W}$ (by Lemma~\ref{lem:uplus-future} and~\ref{lem:purePart-oplus}).
+
+  By definition of $\lrec$, it suffices to prove that 
+  \begin{itemize}
+  \item 
+    $\npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg)$:
     TODO
   \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
-    TODO
+    Since $W_M \future W$, we have seen above that 
+    \begin{equation*}
+      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap[W_M])(W_M)
+    \end{equation*}
+    It suffices to prove that also
+    \begin{itemize}
+    \item $\memSatStack{\emptyset,\emptyset}{\purePart{W_M}}}$
+    \item $\memSatFStack{\ms_\stk,\ms_\stk}{\purePart{W_M}}$
+    \end{itemize}
+    The former follows vacuously.
+    The latter follows by taking $R_\ms(r_a) = (\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})$, $R_W(r_a) = \purePart{W_M}$ for $a \in [\baddr_\stk,\eaddr_\stk]$ by Lemma~\ref{lem:purePart-duplicable} and~\ref{lem:purePart-idempotent} if we can show that:
+    \begin{equation*}
+      \npair{(\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})} \in \stdreg{\{a\},\gc}{\spatialo}.H~ \xi^{-1}(\purePart{W_M})
+    \end{equation*}
+    By definition, it suffices to show that for all $n' < n$, we have that:
+    \begin{equation*}
+      \npair{(\ms_\stk(a),\ms_\stk(a)) \in \lrv(\purePart{W_M})}
+    \end{equation*}
+    But $\ms_\stk(a) = 0$, so this follows easily by definition.
   \item $\npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$:
@@
   \end{itemize}
-
 \end{proof}
 
-\begin{lemma}[Adequacy of logical relation]
+\begin{lemma}[Adequacy of execution configuration LR]
   If

### 2018-06-11 fully-abstract-comp/technical-report.tex
Remove a spurious TODO.
Updated xjump semantics with a missing requirement. Updated memory satisfaction for private stacks to only have addresses greater than the stack base.
Work on the call case of the FTLR proof.
Fix initial execution configuration and finish compat lemma for it.

@@
     \inference{
-      \reg(\pcreg) = c_{\mathrm{main},c} & 
-      \reg(\rdata) = c_{\mathrm{main},d} \\
-      \src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
-      \trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
+      c_{\mathrm{main},c} = \sealed{\sigma_1, c_{\mathrm{main},c}'} &
+      c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'} &
+      \sigma_1 = \sigma_2\\
+      \nonExec{c_{\mathrm{main},d}'}&
+      \reg(\pcreg) = c_{\mathrm{main},c}' & 
+      \reg(\rdata) = c_{\mathrm{main},d}' \\
+      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
+      \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
       \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0\\
@@
   \end{itemize}
-  \item $\overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})$: TODO
+  \item $\overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})$:
 
@@
       \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
-      \npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(\{\rdata\})(W_R)
+      \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
     \end{array}
@@
   \end{itemize}
-  then $\exists W' \future W$ such that
-  \begin{itemize}
-  \item $\npair{(\Phi,\Phi)}\in\lrec(W')$
+  then $\exists W' \future W$ such that for all $n' < n$
+  \begin{itemize}
+  \item $\npair[n']{(\Phi,\Phi)}\in\lrec(W')$
   \end{itemize}
@@
   \begin{itemize}
-  \item $\reg(\pcreg) = c_{\mathrm{main},c}$
-  \item $\reg(\rdata) = c_{\mathrm{main},d}$ 
-  \item $\src{\reg(\rstk) = \stkptr{\rx,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}}$ 
-  \item $\trg{\reg(\rstk) = ((\rx,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)}$ 
+  \item $c_{\mathrm{main},c} = \sealed{\sigma_1, c_{\mathrm{main},c}'}$ 
+  \item $c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'}$
+  \item $\sigma_1 = \sigma_2$
+  \item $\reg(\pcreg) = c_{\mathrm{main},c}'$
+  \item $\reg(\rdata) = c_{\mathrm{main},d}'$ 
+  \item $\nonExec{c_{\mathrm{main},d}'}$
+  \item $\src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}}$ 
+  \item $\trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)}$ 
   \item $\reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0$
@@
   \end{align*}
-  Take $\trust = \trusted$ iff $a \in \ta$. 
   Then $W' = W_M \oplus W_R \oplus W_\pcreg$ (by Lemmas~\ref{lem:purePart-duplicable},~\ref{lem:oplus-distr-uplus},~\ref{lem:oplus-assoc-comm} and~\ref{lem:uplus-assoc-comm}), $W_M \future W$ (by Lemma~\ref{lem:uplus-future}) and $W_\pcreg \future \purePart{W}$ (by Lemma~\ref{lem:uplus-future} and~\ref{lem:purePart-oplus}).
@@
   \item 
-    $\npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg)$:
-    TODO
+    $\npair[n']{( (\reg(\pcreg),\reg(\rdata)), (\reg(\pcreg),\reg(\rdata)) )} \in \lrexj(W_\pcreg)$:
+    Since $W' \future W$, we know from above that:
+    \begin{equation*}
+      \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+    \end{equation*}
+    and we have defined $W_\pc = \purePart{W'}$ and we know that $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq
+    \overline{c_{\mathrm{export}}}$.
+    It follows that 
+    \begin{align*}
+      \npair{(c_{\mathrm{main},c},c_{\mathrm{main},c})} &\in \lrv(W_\pc)\\
+      \npair{(c_{\mathrm{main},d},c_{\mathrm{main},d})} &\in \lrv(W_\pc)
+    \end{align*}
+
+    We also know that $c_{\mathrm{main},c} = \sealed{\sigma, c_{\mathrm{main},c}'}$, $c_{\mathrm{main},d} = \sealed{\sigma, c_{\mathrm{main},d}'}$, $\sigma=\sigma_1 = \sigma_2$, so by definition of $\lrv$, we know that there exist $r_1,r_2 \in \dom(\pwheap[W_\pc])$ such that $\pwheap[W_\pc](r_1) = (\pure,\_,H_{\sigma,1})$, $\pwheap[W_\pc](r_2) = (\pure,\_,H_{\sigma,2})$ and
+    \begin{align*}
+      \npair[n']{(c_{\mathrm{main},d}',c_{\mathrm{main},d}')} &\in H_{\sigma,1}~ \sigma~ \xi^{-1}(W_\pc) \text{ for all } n' < n\\
+      \npair[n']{(c_{\mathrm{main},d}',c_{\mathrm{main},d}')} &\in H_{\sigma,2}~ \sigma~ \xi^{-1}(W_\pc) \text{ for all } n' < n
+    \end{align*}
+    By definition of $W_\stk$, we know that it contains no regions that define $H_\sigma$.
+    Therefore, using the definition of $W'$, we can deduce that $r_1,r_2 \in \dom(\pwheap)$.
+    From the fact that 
+    \begin{equation*}
+      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
+    \end{equation*}
+    it then follows that $\sigma$ can only be defined in $H_\sigma$ for only one region, so that $r_1 = r_2 = r$ and $H_{\sigma,1} = H_{\sigma,2} = H_\sigma$.
+
+    Since $W_\pc = \purePart{W'}$, we have (in both the linear and non-linear case), that for $n' < n$, $\npair[n']{( (c_{\mathrm{main},c},c_{\mathrm{main},d}), (c_{\mathrm{main},c},c_{\mathrm{main},d}) )} \in \lrexj(W_\pcreg)$.
+    Since $\reg(\pcreg) = c_{\mathrm{main},c}'$ and $\reg(\rdata) = c_{\mathrm{main},d}'$, this is what we set out to prove.
+
   \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
@@
     \end{equation*}
+    and hence by Lemma~\ref{lem:downwards-closed}:
+    \begin{equation*}
+      \npair[n']{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap[W_M])(W_M)
+    \end{equation*}
     It suffices to prove that also
     \begin{itemize}
-    \item $\memSatStack{\emptyset,\emptyset}{\purePart{W_M}}}$
-    \item $\memSatFStack{\ms_\stk,\ms_\stk}{\purePart{W_M}}$
+    \item $\memSatStack[n']{\emptyset,\emptyset}{\purePart{W_M}}}$
+    \item $\memSatFStack[n']{\ms_\stk,\ms_\stk}{\purePart{W_M}}$
     \end{itemize}
@@
     \begin{equation*}
-      \npair{(\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})} \in \stdreg{\{a\},\gc}{\spatialo}.H~ \xi^{-1}(\purePart{W_M})
-    \end{equation*}
-    By definition, it suffices to show that for all $n' < n$, we have that:
-    \begin{equation*}
-      \npair{(\ms_\stk(a),\ms_\stk(a)) \in \lrv(\purePart{W_M})}
+      \npair[n']{(\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})} \in \stdreg{\{a\},\gc}{\spatialo}.H~ \xi^{-1}(\purePart{W_M})
+    \end{equation*}
+    By definition, it suffices to show that for all $n'' < n'$, we have that:
+    \begin{equation*}
+      \npair[n'']{(\ms_\stk(a),\ms_\stk(a)) \in \lrv(\purePart{W_M})}
     \end{equation*}
     But $\ms_\stk(a) = 0$, so this follows easily by definition.
-  \item $\npair{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$:
-    TODO
+
+  \item $\npair[n']{(\reg,\reg)} \in \lrr(\{\rdata\})(W_R)$:
+
+    We have that $\reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0$, so by definition, it suffices to prove that
+    \begin{equation*}
+      \npair[n']{(\reg(r_\stk),\reg(r_\stk))} \in \lrv(W_R)
+    \end{equation*}
+
+    Since $\src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}}$ and $\trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)}$, it suffices to prove (by definition) that
+    \begin{align*}
+      \npair[n']{[b_\stk,e_\stk]} &\in \stackReadCond{W_R}\\
+      \npair[n']{[b_\stk,e_\stk]} &\in \stackWriteCond{W_R}
+    \end{align*}
+    For both, we can take $S = \dom(\pwfree[W_\stk]) = \{r_a \mid a \in [\baddr_\stk,\eaddr_\stk]\}$, $R(r_\addr) = \{\addr\}$, and then it suffices to prove that for all $r_\addr$, $\pwfree[W_R](r_\addr).H \nsubeq \stdreg{\{\addr\},\gc}{\spatialo}$ resp.
+$\pwfree[W_R](r_\addr).H \nsupeq \stdreg{\{\addr\},\gc}{\spatialo}$.
+    Both follow easily since 
+    $\pwfree[W_R](r_a) = \pwfree[W_\stk](r_a) = \stdreg{\{a\},\gc}{\spatialo}$.
   \end{itemize}

### 2018-06-11 fully-abstract-comp/ftlr-proof-cases.tex

@@
   \updReg{\rretd}{\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}} \\
-  \updReg{\rretc}{\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\var{off\_return})}} \\
+  \updReg{\rretc}{\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\retoffset)}} \\
   \updReg{\rtmp{1}}{0}\\
@@
 \]
-where $\Phi_T(r_i) = \sealed{\sigma,c_i'} \text{ for $i\in\{1,2\}$}$ and $\var{off\_return} = 15$ which is the offset to the return code.
+where $\Phi_T(r_i) = \sealed{\sigma,c_i'} \text{ for $i\in\{1,2\}$}$ and $\retoffset = 15$ which is the offset to the return code.
 Now using Lemma~\ref{lem:lro-anti-red-gen} again, it suffices to show
@@
   where $W_{R,\rretd}'$ is the part of $W_R'$ with ownership over $r_\var{priv\_stk}$.
-  From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}$.
+  From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\sigrets',\_,M(r),\gc}$.
   Further by $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one
@@
   That is $\mscode'$ contains the call we are considering.
-  \lau{TODO: Make the above more precise and possibly put into a lemma.}
-  Use the $r$ in question as the witness. We now have to show:
+  \lau{TODO: Make the above more precise and possibly put into a lemma (also needed in the next case).}
+  Use the $r$ in question as the witness.
+  The $\xReadCond{}$ gives us $\pwheap(r) \nequal \codereg{\_,\_,\mscode',\gc}$.
+  As it is a pure region, it is also present in the future world we consider.
+  We now have to show:
   \begin{enumerate}[label=\alph*)]
@@
 
-\item[Case $\rretc$:] \lau{TODO}
+\item[Case $\rretc$:]
+  We have to show 
+  \[
+    \npair[n'-1]{\left(\array{l}\sealed{\sigma',\retptrc(\baddr,\eaddr,\aaddr +
+        \calllen)},\\\sealed{\sigma',((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)}\endarray\right)} \in \lrv(W_{R,\rretc}')
+  \]
+  where $W_{R,\rretc}'$ has the ownership of $W$.
+  Just as in the previous case, we know that for some region $r$ there exists $\mscode' \subseteq \mscode$ such that: $\pwheap(r) \nequal \codereg{\sigrets',\_,\mscode',\gc}$ where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
+  It follows easily from the definition of $H_\sigma^\var{code}$ that
+  \begin{equation}
+    \label{eq:code-caps-in-h-sigma}
+    \npair[n'']{\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr + \calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)\endarray\right)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_{R,\rretc}')
+  \end{equation}
+  for $n'' < n'-1$.
+  Both capabilities are non-linear, so $\isLinear{\retptrc(\baddr,\eaddr,\aaddr + \calllen)}$ iff $\isLinear{((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)}$ is indeed the case.
+  Finally we need to show:
+  \[
+    \begin{array}{l}
+      \forall W'' \future \purePart{W_{R,\rretc}'}, W_o, n'' < n'-1,\\
+      \quad  \npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \npair[n'']{\retptrc(\baddr,\eaddr,\aaddr + \calllen),\vsc_S',((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset),\vsc_T'} \in \lrexj(W''\oplus W_o))
+    \end{array}
+  \]
+  To this end let $W'' \future \purePart{W_{R,\rretc}'}$ and $W_o$ be given s.t.
+  $W'' \future \purePart{W_{R,\rretc}'} \oplus W_o$ is defined. Further, let
+  $\npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \;
+  \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \;
+  \xi^{-1}(W_o)$ be given and show 
+  \[
+    \npair[n'']{\retptrc(\baddr,\eaddr,\aaddr + \calllen),\vsc_S',((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset),\vsc_T'} \in \lrexj(W''\oplus W_o))
+  \]
+  Now let $n''' \leq n''$ be given along with $\reg[3]_S$, $\reg[3]_T$, $\ms^{(3)}_S$, $\ms^{(3)}_T$,$\ms^{(3)}_\stk$, $\stk^{(3)}$, $W_R''$, and $W_M''$ such that
+  \begin{itemize}
+  \item $W'' \oplus W_o \oplus W_R'' \oplus W_M''$ is defined
+  \item $\memSat[n''']{\ms^{(3)}_S,\ms^{(3)}_\stk, \stk^{(3)}, \ms^{(3)}_T}{W_M''}$
+  \item $\npair[n''']{(\reg[3]_S,\reg[3]_T)} \in \lrr(W_R'')$
+  \end{itemize}
+
+  Based on $H_\sigma^\var{code}$, there are three possible values for $\vsc_S'$ and $\vsc_T'$.
+  In the first case, $\vsc_S'$ is a $\retptrc$ and $\vsc_T'$ is a capability with permission $\rx$.
+  In this case, $\xjumpResult{}{}{}$ will give to failed configurations which is trivially in the observation relation.
+  In the next case, it is required that $\sigma' \in \sigcloss'$, but this cannot be the case as \lau{TODO I think this case should follow from $H^\var{code}$ and the comp-code judgement, but it does not.
+    Specifically, the comp-code judgement (for now) does not require $\sigrets'$ and $\sigcloss'$ to be disjointed.}
+
+  This leaves us with one final case, namely $\vsc_S'= \retptrd(\baddr_\stk',\eaddr_\stk')$ and $\vsc_T'=((\rw,\linear),\baddr_\stk',\eaddr_\stk',\baddr_\stk'-1)$.
+  Further we know
+  \[
+    \begin{array}{l}
+    \exists r \in \addressable{\linear,\pwpriv[W_o]} \ldotp \pwpriv[W_o](r).H \nequal (\stareg[(\ms_{\priv,S}^{(3)},\ms_{\priv,T}^{(3)}),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+    \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr_\stk',\eaddr_\stk'] \tand\\
+    \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+    \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets 
+    \end{array}
+  \]
+  Call this region $r'$.
+  By the above, the fact that the code capability pair is in $H_\sigma^\var{code}$ (\ref{eq:code-caps-in-h-sigma}) and Lemma~\ref{lem:unique-ret-seals}, we get $\aaddr' = \aaddr$.
+  This means that $W_M''$ and $W_R''$ has this region.
+
+  We know that the two register-files are related which in particular means that
+  the values in register $\rstk$ are related. Now consider the following cases:
+  \lau{TODO: Perhaps make a lemma that says if source fail and target step to
+    fail, then the two configurations are in $\lro$.}
+  \begin{itemize}
+  \item $\Phi_S(\rstk) \neq \stkptr{\_,\_,\_,\_}$\\
+    In this case due to $\reg[3]_S$ being related to $\reg[3]_T$, there are three cases we need to consider.
+    In all cases, the source configuration will fail because the value in the stack register is not a stack capability.
+    If we can argue that the target configuration also will fail, then the two are in the observation relation.
+    First, if $\reg[3]_T = \sealed{\sigma_\var{ret\_\stk},\vsc_T''}$, then the return code will fail when the base address (a sealed capability has no base address, so the instruction returns $-1$) is compared to $\stkb$.
+    Second, if $\reg[3]_T = \seal{\_,\_,\_}$, then the target execution fails when it attempts to splice this seal with $\vsc_T'$.
+    Finally, $\reg[3]_T = ((\perm_\var{ret\_\stk},\lin_\var{ret\_\stk}),\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_)$ and $\npair[n''']{[\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk}]} \in \readCond{\lin_\var{ret\_\stk},W_{R,\rstk}''}$ is satisfied.
+    This means that it is satisfied by some heap region, but by the memory satisfaction assumption $\stkb$ must be in the free stack part of the world.
+    This means that the execution will fail that $\stkb$ check.
+  \item $\Phi_S(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\_,\_}$ and $\baddr_\var{ret\_\stk} \neq \stkb$\\
+    Here the source side will fail the xjmp as the base address is not $\stkb$.
+    Similarly on the target side, the return code will fail the $\stkb$ check.
+  \item $\Phi_S(\rstk) = \stkptr{\_,\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and either $\eaddr_\var{ret\_\stk} +1 \neq \baddr_\stk' $ or $\perm_\var{ret\_\stk} \neq \rw$ or
+    $\baddr_\var{ret\_\stk} > \eaddr_\var{ret\_\stk}$.\\
+    In this case, the source configuration will fail as one of the conditions in $\var{xjumpResult}$ will not be met.
+    On the target side, the splice will fail as either the two capabilities being spliced don't line up, the permissions don't match, or the range of authority is empty, respectively.
+  \item $\Phi_S(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and $\eaddr_\var{ret\_\stk} +1 = \baddr_\stk' $ and $\perm_\var{ret\_\stk} = \rw$ and $\baddr_\var{ret\_\stk} \leq \eaddr_\var{ret\_\stk}$.
+    \\
+    We would like to show that $\ms_{\priv,S}^{(3)}$ is the top most stack frame and that $r'$ governs it.
+    By the memory satisfaction assumption and the presence of $r'$ in $W_M''$ we know that $\stk^{(3)}$ is non-empty.
+    By the memory satisfaction on the private stack, the following must be the case:
+    \[
+      \begin{array}{l}
+        \stk^{(3)} = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
+        \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
+        \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a')\\
+      \end{array}
+    \]
+    Assume for contradiction $\ms_{\priv,S}^{(3)}$ is not the top frame. In that
+    case $\dom(\ms_0) \neq \emptyset$ and $\forall a \in \dom(\ms_0) \ldotp
+    \ldotp \stkb < a < b_\stk$ at the same time, we know
+    \[
+      \npair[n'']{(\stkptr{\rw,\stkb,\eaddr_\var{ret\_\stk},\_},((\rw,\linear),\stkb,\eaddr_\var{ret\_\stk},\_))}
+      \in \lrv(W_{R,\rstk}'')
+    \]
+    which means that the free stack part of the world contains a region that at least governs $[\stkb,\eaddr_\var{ret\_\stk}]$.
+    Combine this with $\eaddr_\var{ret\_\stk} +1 = \baddr_\stk' $, we can conclude that no such address can exist in $\ms_0$, so it must be empty, but this cannot be the case either.
+    Therefore, the top stack frame must contain $\ms_{\priv,S}^{(3)}$.
+    Further, due to the disjointedness required by memory satisfaction, it must be $r'$ that governs this stack frame.
+    This also means that we have $\opc_0 =  \aaddr+\calllen$.
+    With this, we have all the requirements for $\var{xjumpResult}$ satisfied on
+    both sides which allows us to pick the necessary configurations:
+    \begin{multline*}
+      \Phi^{(4)}_S=\xjumpResult{r_1}{r_2}{(\ms_S^{(3)},\reg[3]_S,(\ms_{\priv,S}^{(3)},\aaddr+\calllen)
+        :: \stk^{(3)}_\var{rest},\ms_\stk^{(3)})} =\\
+      (\ms_S^{(3)},\reg[3]_S
+      \arraycolsep=0pt
+      \begin{array}[t]{l}
+        \updReg{\pcreg}{((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen)} \\
+        \updReg{\rdata}{0} \\
+        \updReg{\rstk}{\stkptr{\rw,\stkb,e_\stk',\eaddr_\var{ret\_stk}+1}} \\
+        \updReg{\rtmp{1}}{0} \\
+        \updReg{\rtmp{2}}{0}
+      \end{array}
+      ,\stk^{(3)}_\var{rest},\ms_{\priv,S}^{(3)} \uplus \ms_\stk^{(3)})
+    \end{multline*}
+    and
+    \begin{multline*}
+      \Phi^{(4)}_T=\xjumpResult{r_1}{r_2}{(\ms_T^{(3)},\reg[3]_T)} =\\
+      (\ms_T^{(3)},\reg[3]_S
+      \arraycolsep=0pt
+      \begin{array}[t]{l}
+        \updReg{\pcreg}{((\rx,\normal),\baddr,\eaddr,\aaddr+\retoffset)} \\
+        \updReg{\rdata}{((\rw,\linear),\stkb,\eaddr_\var{ret\_stk})}
+      \end{array})
+    \end{multline*}
+    It now remains to show
+    \[
+      \npair[n''']{(\Phi^{(4)}_S,\Phi^{(4)}_T)} \in \lro
+    \]
+    Use Lemma~\ref{lem:lro-anti-red-gen} by which it suffices to show the
+    following two things:
+    \begin{itemize}
+    \item
+      \[
+        \Phi^{(4)}_T \step[l] \Phi^{(5)}_T = (\ms_T^{(3)},\reg[3]_S
+        \arraycolsep=0pt
+        \begin{array}[t]{l}
+          \updReg{\pcreg}{((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen)} \\
+          \updReg{\rdata}{0} \\
+          \updReg{\rstk}{((\rw,\linear),\stkb,e_\stk',\eaddr_\var{ret\_stk}+1)} \\
+          \updReg{\rtmp{1}}{0} \\
+          \updReg{\rtmp{2}}{0}
+        \end{array})
+      \]
+      for some number of steps $l$. This follows immediately from the
+      operational semantics.
+    \item
+      \[
+        \npair[n''']{(\Phi^{(4)}_S,\Phi^{(5)}_T)} \in \lro
+      \]
+      Here we need to use the induction hypothesis.
+    \end{itemize}
+
+  \end{itemize}
+
 \end{description}
+This concludes case \ref{case:ftlr:scall:reg}
 
@@
 
-\begin{lemma}
-
-  If $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ and
-  $\memSat{\ms_S \uplus \mscode, \stk,\ms_\stk,\ms_T}{W}$
-  $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and
-  \[
-    \begin{array}{l}
-      \exists r \in \dom(\pwheap) ,\mscode' \ldotp \\
-      \quad \pwheap(r) = \codereg{\sigrets',\sigcloss',\mscode',(\ta,\stkb)} \text{ and } \\
-      \quad [\baddr,\eaddr] \subseteq \dom(\mscode') \text{ and } \\
-      \quad \sigma \in \sigrets' \subseteq \sigrets \text{ and } \\
-      \quad \sigclos \subseteq \sigclos'
-    \end{array}
-  \]
-\end{lemma}
-
 \end{proof}
 
+\begin{lemma}[Unique return seals]
+  \label{lem:unique-ret-seals}
+  If
+  \begin{itemize}
+  \item $\mscode([\aaddr..\aaddr+\calllen-1]) = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$
+  \item $\mscode([\aaddr'..\aaddr'+\calllen-1]) = \overline{\scall{\offpc',\offsigma'}{r_1'}{r_2'}}$
+  \item $\mscode(\aaddr + \offpc) = \seal{\sigma_b,\sigma_e,\sigma_b}$ and $\sigma = \sigma_b + \offpc$
+  \item $\mscode(\aaddr' + \offpc') = \seal{\sigma_b',\sigma_e',\sigma_b'}$ and $\sigma = \sigma_b' + \offpc'$
+  \end{itemize}
+  then
+  \[
+    \aaddr = \aaddr'
+  \]
+  and $\offpc = \offpc'$ and $\offsigma = \offsigma'$ and $r_1 = r_1'$ and $r_2 = r_2'$ and $\sigma_b = \sigma_b'$ and $\sigma_e = \sigma_e'$.
+\end{lemma}
+\begin{proof}
+  \lau{TODO}
+\end{proof}
 %%% Local Variables:

### 2018-06-12 fully-abstract-comp/technical-report.tex
Fix latex compilation.
Some more details on one of the remaining TODOs.

@@
         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
+        \stkb \leq \eaddr_\stk \tand\\
         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
@@
       \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
-      \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp a < a') \wedge\\
+      \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
       \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
@@
         \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
-                                  \npair{[\baddr,\eaddr]} \in \xReadCond{W}
-                                  \lin = \normal \wedge \\ \endarray
+        \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
+                                  \lin = \normal \\ \endarray
       \end{array}
@@
         \aaddr = \aaddr' + \retoffset \tand \\
-        \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
+        \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
         \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
@@
         \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\_} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
+        \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
       \end{array}
@@
 
-     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs} that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(W)$, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs}, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
 
@@
      \end{equation*}
+
+     Since $\sigma\in\sigcloss$, it follows from $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$ that
+     \begin{equation*}
+       \npair[n']{(\vsc_S',\vsc'_T)} \in \lrvg{\trust}(W_o)
+     \end{equation*}
+     with $\trust = \untrusted$ iff $\dom(\code \mathrel{\#} \ta)$ and $\trust = \trusted$ iff $\dom(\code) \subseteq \ta$.
+
+     As above, it follows from Lemma~\ref{lem:ftlr-component-code-ptrs}, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$ that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(\purePart{W})$.
+
      \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
@@
     \begin{itemize}
-    \item $\memSatStack[n']{\emptyset,\emptyset}{\purePart{W_M}}}$
+    \item $\memSatStack[n']{\emptyset,\emptyset}{\purePart{W_M}}$
     \item $\memSatFStack[n']{\ms_\stk,\ms_\stk}{\purePart{W_M}}$

### 2018-06-12 fully-abstract-comp/technical-report.tex
Finish another proof.

@@
       \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
+      \reg_S(\pcreg) \neq \retptrc(\_) \wedge 
+      \reg_S(\rdata) \neq \retptrd(\_) \wedge 
+      \nonExec{\reg_S(\rdata)} \wedge
+      \nonExec{\reg_T(\rdata)}\\
       \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
@@
 
+  \item $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$, $\nonExec{\reg_S(\rdata)}$ and $\nonExec{\reg_T(\rdata)}$:
+    This follows from the facts that $\reg(\pcreg) = c_{\mathrm{main},c}'$, $\reg(\rdata) = c_{\mathrm{main},d}'$, 
+    $\nonExec{c_{\mathrm{main},d}'}$, $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and the fact that $\overline{c_{\mathrm{export}}}$ are also valid
+
   \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
@@
 \end{lemma}
+\begin{proof}
+  First, assume that 
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$
+  \item $i \leq n$
+  \item $\Phi_S \sterm[i]{\gc}$
+  \end{itemize}
+
+  Assume w.l.o.g. that $\Phi_S = (\ms_S,\reg_S,\stk,\ms_\stk)$, $\Phi_T = (\ms_T,\reg_T)$ and $\gc = (\ta,\stkb)$.
+  Then it follows from $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$ that there exist $W_M,W_R,W_\pcreg$ such that
+  \begin{itemize}
+  \item $W = W_M \oplus W_R \oplus W_\pcreg$
+  \item $\npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj[\preceq,\gc](W_\pcreg)$
+  \item $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$, $\nonExec{\reg_S(\rdata)}$ and $\nonExec{\reg_T(\rdata)}$:
+  \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$
+  \item $\npair{\stpair{\reg}{\reg}} \in \lrr[\preceq,\gc](\{\rdata\})(W_R)$
+  \end{itemize}
+
+  We can then instantiate the conditions from $\lrexj$ with the other conditions from this list to obtain $\Phi_S',\Phi_T'$ such that
+  \begin{itemize}
+  \item $\Phi_S' = \xjumpResult{\reg_S(\pcreg)}{\reg_S(\rdata)}{\Phi_S}$ 
+  \item $\Phi_T' = \xjumpResult{\reg_T(\pcreg)}{\reg_T(\rdata)}{\Phi_T}$
+  \item $\npair{\left(\Phi_S', \Phi_T' \right)}\in \lrol$
+  \end{itemize}
+
+  Using the facts that $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$ and $\nonExec{\reg_S(\rdata)}$, we know (by definition of $\xjumpResult{\_}{\_}{\_}$) that  
+  \begin{itemize}
+  \item $\Phi_S' = \xjumpResult{\reg_S(\pcreg)}{\reg_S(\rdata)}{\Phi_S} = \Phi_S\updReg{\pcreg}{\reg_S(\pcreg)}\updReg{\rdata}{\reg_S(\rdata)} = \Phi_S$ 
+  \item $\Phi_T' = \xjumpResult{\reg_T(\pcreg)}{\reg_T(\rdata)}{\Phi_T} = \Phi_T\updReg{\pcreg}{\reg_T(\pcreg)}\updReg{\rdata}{\reg_T(\rdata)} = \Phi_T$ 
+  \end{itemize}
+
+  From $\npair{\left(\Phi_S, \Phi_T \right)}\in \lrol$, it follows immediately that if $\Phi_S \sterm[i]{\ta,\stkb}$ with $i \le n$, then $\Phi_T \term$ as required.
+
+  The proof in the other direction is directly analogous.
+\end{proof}
 

### 2018-06-12 fully-abstract-comp/technical-report.tex
Finish a consequence lemma.

@@
 \begin{lemma}[Compatibility lemma for linking]
+  \label{lem:compat-linking}
   If
@@
 \begin{lemma}[Compatibility lemma for initial execution configuration construction]
+  \label{lem:compat-initial-ec}
   If
@@
   If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
-  $\npair{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \oplus W_2)$.
+  $\npair{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$.
 \end{lemma}
+\begin{proof}
+  By definition, we have that $C_S \bowtie P_S \rightsquigarrow C_S[P_S]$ and $C_T \bowtie P_T \rightsquigarrow C_T[P_T]$.
+
+  The result follows directly from Lemmas~\ref{lem:compat-linking} and~\ref{lem:compat-initial-ec}.
+\end{proof}
 

### 2018-06-12 fully-abstract-comp/technical-report.tex
Fix typo.

@@
 \begin{definition}
-  A component $\var{comp})$ with $\dom(\var{comp}.\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
+  A component $\var{comp}$ with $\dom(\var{comp}.\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
   \begin{itemize}

### 2018-06-12 fully-abstract-comp/technical-report.tex
typo.

@@
   \item $\plug{\context}{\var{comp}}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\plug{\context}{\var{comp}}$ handle trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
+  \item $\plug{\context}{\var{comp}}$ handles trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
   \item $\plug{\context}{\var{comp}}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).

### 2018-06-12 fully-abstract-comp/technical-report.tex
Added disjointedness requirement on ret seals and clos seals to comp-code judgement.

--- 2018-06-12 fully-abstract-comp/technical-report.tex
@@
     \text{$\mscode$ has no hidden calls}\\
+    \sigrets \mathrel{\#} \sigcloss \\
     \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
@@
 \begin{lemma}[Compatibility lemma for linking]
-  \label{lem:compat-linking}
   If
@@
       \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
-      \reg_S(\pcreg) \neq \retptrc(\_) \wedge 
-      \reg_S(\rdata) \neq \retptrd(\_) \wedge 
-      \nonExec{\reg_S(\rdata)} \wedge
-      \nonExec{\reg_T(\rdata)}\\
       \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
@@
 \begin{lemma}[Compatibility lemma for initial execution configuration construction]
-  \label{lem:compat-initial-ec}
   If
@@
 
-  \item $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$, $\nonExec{\reg_S(\rdata)}$ and $\nonExec{\reg_T(\rdata)}$:
-    This follows from the facts that $\reg(\pcreg) = c_{\mathrm{main},c}'$, $\reg(\rdata) = c_{\mathrm{main},d}'$, 
-    $\nonExec{c_{\mathrm{main},d}'}$, $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and the fact that $\overline{c_{\mathrm{export}}}$ are also valid
-
   \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
@@
 \end{lemma}
-\begin{proof}
-  First, assume that 
-  \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$
-  \item $i \leq n$
-  \item $\Phi_S \sterm[i]{\gc}$
-  \end{itemize}
-
-  Assume w.l.o.g. that $\Phi_S = (\ms_S,\reg_S,\stk,\ms_\stk)$, $\Phi_T = (\ms_T,\reg_T)$ and $\gc = (\ta,\stkb)$.
-  Then it follows from $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$ that there exist $W_M,W_R,W_\pcreg$ such that
-  \begin{itemize}
-  \item $W = W_M \oplus W_R \oplus W_\pcreg$
-  \item $\npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj[\preceq,\gc](W_\pcreg)$
-  \item $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$, $\nonExec{\reg_S(\rdata)}$ and $\nonExec{\reg_T(\rdata)}$:
-  \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$
-  \item $\npair{\stpair{\reg}{\reg}} \in \lrr[\preceq,\gc](\{\rdata\})(W_R)$
-  \end{itemize}
-
-  We can then instantiate the conditions from $\lrexj$ with the other conditions from this list to obtain $\Phi_S',\Phi_T'$ such that
-  \begin{itemize}
-  \item $\Phi_S' = \xjumpResult{\reg_S(\pcreg)}{\reg_S(\rdata)}{\Phi_S}$ 
-  \item $\Phi_T' = \xjumpResult{\reg_T(\pcreg)}{\reg_T(\rdata)}{\Phi_T}$
-  \item $\npair{\left(\Phi_S', \Phi_T' \right)}\in \lrol$
-  \end{itemize}
-
-  Using the facts that $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$ and $\nonExec{\reg_S(\rdata)}$, we know (by definition of $\xjumpResult{\_}{\_}{\_}$) that  
-  \begin{itemize}
-  \item $\Phi_S' = \xjumpResult{\reg_S(\pcreg)}{\reg_S(\rdata)}{\Phi_S} = \Phi_S\updReg{\pcreg}{\reg_S(\pcreg)}\updReg{\rdata}{\reg_S(\rdata)} = \Phi_S$ 
-  \item $\Phi_T' = \xjumpResult{\reg_T(\pcreg)}{\reg_T(\rdata)}{\Phi_T} = \Phi_T\updReg{\pcreg}{\reg_T(\pcreg)}\updReg{\rdata}{\reg_T(\rdata)} = \Phi_T$ 
-  \end{itemize}
-
-  From $\npair{\left(\Phi_S, \Phi_T \right)}\in \lrol$, it follows immediately that if $\Phi_S \sterm[i]{\ta,\stkb}$ with $i \le n$, then $\Phi_T \term$ as required.
-
-  The proof in the other direction is directly analogous.
-\end{proof}
 
@@
   If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
-  $\npair{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$.
+  $\npair{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \oplus W_2)$.
 \end{lemma}
-\begin{proof}
-  By definition, we have that $C_S \bowtie P_S \rightsquigarrow C_S[P_S]$ and $C_T \bowtie P_T \rightsquigarrow C_T[P_T]$.
-
-  The result follows directly from Lemmas~\ref{lem:compat-linking} and~\ref{lem:compat-initial-ec}.
-\end{proof}
 

### 2018-06-12 fully-abstract-comp/ftlr-proof-cases.tex
Fix typos.
Add some extra non-expansiveness statements.
typo
Added some argumentation for the form of r_1 and r_2 not the same register and not pc.

@@
     \item $\nonExec{c_2}$
-    \item $\Phi_S(\stk) = \stkptr{\rw, \baddr_\stk, \eaddr_\stk, \aaddr_\stk}$
+    \item $\Phi_S(\rstk) = \stkptr{\rw, \baddr_\stk, \eaddr_\stk, \aaddr_\stk}$
     \item $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$
@@
 \\\\
-In case \ref{case:ftlr:call}, we first need to let the target execution catch
-up. That is $\Phi_T' \nstep[15]{} \Phi_T''$ for 
+\lau{TODO: look through this case and find the place where the ''no seal cross boundary thing'' should have been used.}
+In case \ref{case:ftlr:call}, we may assume $r_1 \neq r_2$\footnote{If the register contains a data capability, then the execution fails in the step after the jump.
+  If it is an executable capability, then the xjump fails as it does not permit executable capabilities for the data part.}
+and $r_i \neq \pcreg$ for $i \in \{1,2\}$\footnote{The pc is executable which causes the xjump to fail.}
+as this will cause the execution to fail.
+We need to let the target execution catch up.
+That is $\Phi_T' \nstep[15]{} \Phi_T''$ for
 \[

### 2018-06-12 fully-abstract-comp/ftlr-proof-cases.tex
Updated semantics of scall, so it matches what the implementation does when one of the registers is the temp register.
typo

@@
     \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
+    \item $r_1 \neq \rtmp{1}$ and $r_2 \neq \rtmp{1}$
     \item $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$
@@
   If it is an executable capability, then the xjump fails as it does not permit executable capabilities for the data part.}
-and $r_i \neq \pcreg$ for $i \in \{1,2\}$\footnote{The pc is executable which causes the xjump to fail.}
+and $r_i \neq \pcreg$ for $i \in \{1,2\}$\footnote{The pc is executable which causes the xjump to fail.}.
 as this will cause the execution to fail.

### 2018-06-12 fully-abstract-comp/ftlr-proof-cases.tex
Small text update

@@
 
-Case \ref{case:ftlr:scall:reg}, using assumption $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
-We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership.
+Case \ref{case:ftlr:scall:reg}: We need to split the ownership of $W_R'$.
+From assumption $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
+We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership (i.e.\ $\purePart{W_{R,1}'}$).
 region $\rstk$ maps to a world with the same ownership, but of course without the now revoked regions.

### 2018-06-12 fully-abstract-comp/ftlr-proof-cases.tex
Discharged the remaining cases for showing register file safe in call case of FTLR.

@@
 \begin{description}
-\item[Case $\rstk$:] \lau{TODO}
-\item[Case $\rtmp{1}$:] \lau{TODO}
-\item[Case $r_1$,$r_2$:] \lau{TODO}
+\item[Case $\rstk$:] Show:
+  \[
+    \npair[n'-1]{(\stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1},((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1))} \in \lrv(W_{R,\rstk}')
+  \]
+  We know by Assumption~\ref{item:reg-ass}, we know
+  \[
+    \npair[n']{(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk},((\rw, \linear), \baddr_\stk ,\eaddr_\stk, \aaddr_\stk)} \in \lrrg{\trust}(W_{R,\stk})
+  \]
+  which by the $\stackReadCond{}$ gives $S \subseteq
+  \addressable{\linear,\pwfree}$ and $R : S \fun \powerset{\nats}$.
+  We need to pick an $S'$ to argue
+  \[
+    \npair[n'-1]{[\baddr_\stk,\aaddr_\stk-1]} \in \stackReadCond{W_{R,\rstk}'}
+  \]
+  To this end pick $S' = R^{-1}(\left(\bigcup_{r\in S} R(r) \right) \setminus [\aaddr_\stk,\eaddr_\stk])$ and $R'$ to be $R$ limited to $S'$.
+  As we exclude all the revoked regions and the ownership otherwise remains the
+  same in $W_{R,\rstk}'$ as in $W_{R,\rstk}$, the regions in $S'$ are exactly the same in the new world as they were in $W_{R,\rstk}$.
+  So what we need to show follows immediately from
+  \[
+    \npair[n']([\baddr_\stk,\eaddr_\stk]) \in \stackReadCond{W_{R,\rstk}}
+  \]
+  and by Lemma~\ref{lem:downwards-closed}.
+
+  We show
+  \[
+    \npair[n'-1]{[\baddr_\stk,\aaddr_\stk-1]} \in \stackWriteCond{W_{R,\rstk}'}
+  \]
+  in the same way.
+
+\item[Case $\rtmp{1}$:] Show:
+  \[
+    \npair[n'-1]{(0,0)} \in \lrv(W_{R,\rtmp{1}}')
+  \]
+  Follows immediately from the definition.
+
+\item[Case $r_1$,$r_2$:] The two cases are symmetric, so we just show the $r_1$ case:
+  \[
+    \npair[n'-1]{(\linCons{c_1},\linCons{\Phi_T(r_1)})} \in \lrv(W_{R,r_1}')
+  \]
+  Follows from the related register files assumption and Lemma~\ref{lem:lincons-lrv}.
+
 \item[Case $\rretd$:] We have to show 
@@
 \end{proof}
+
+\begin{lemma}
+  \label{lem:lrv-relates-linearity}
+  For all $W$, $n$, $w_1$, $w_2$ if
+  \begin{itemize}
+  \item $\npair{(w_1,w_2)} \in \lrvg{\trust}(W)$
+  \end{itemize}
+  then
+  \[
+    \isLinear{w_1} \text{ iff } \isLinear{w_2}
+  \]
+\end{lemma}
+\begin{proof}
+  (Trivial) We consider the possible cases for $w_1$ and $w_2$
+  \begin{itemize}
+  \item Case $w_1,w_2 \in \ints$: By definition of $\isLinear{}$.
+  \item Case $w_1 = \sealed{\sigma,\vsc_1}$ and $w_2 = \sealed{\sigma,\vsc_2}$:
+    From the assumption $\npair[n]{(w_1,w_2)} \in \lrv(W)$ we get the desired result.
+  \item Case $w_1 = \seal{\_,\_,\_}$ and $w_2 = \seal{\_,\_,\_}$: By definition of $\isLinear{}$.
+  \item Case $w_1 = \stkptr{\_,\_,\_,\_}$ and $w_2 = ((\_,\linear),\_,\_,\_)$: By definition of $\isLinear{}$, stack pointers are linear.
+  \item Case $w_1 = ((\_,\lin),\_,\_,\_)$ and $w_2 = ((\_,\lin),\_,\_,\_)$: By definition of $\isLinear{}$.
+  \end{itemize}
+  If $\trust = \trusted$, then there are two more cases to consider, but like the above cases, they are trivial.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:lincons-lrv}
+  For all $W' \future W$, $n$, $w_1$, $w_2$
+   if
+    \begin{itemize}
+    \item $\isLinear{w_1}$ or $\isLinear{w_2}$
+    \end{itemize} or
+    \begin{itemize}
+    \item $\npair{(w_1,w_2)} \in \lrv(W)$
+    \item $\nonLinear{w_1}$ or $\nonLinear{w_2}$
+    \end{itemize}
+    then
+    \[
+      \npair{(\linCons{w_1},\linCons{w_2})} \in \lrv(\purePart{W'})
+    \]
+\end{lemma}
+\begin{proof}
+  For the first set of assumptions, we can conclude
+  $\isLinear{w_1}$ and $\isLinear{w_2}$ by Lemma~\ref{lem:lrv-relates-linearity} which means that we need to argue
+  \[
+    \npair{(0,0)} \in \lrv(\purePart{W'})
+  \]
+  which is trivially true.
+
+  For the second set of assumptions, we can conclude $\nonLinear{w_1}$ and
+  $\nonLinear{w_2}$ by Lemma~\ref{lem:lrv-relates-linearity} which means that we
+  need to show $\npair{(w_1,w_2)} \in \lrv(\purePart{W'})$ which is true by
+  assumption and Lemma~\ref{lem:monotonicity} and \ref{lem:non-linear-pure}.
+\end{proof}
 %%% Local Variables:

### 2018-06-12 fully-abstract-comp/ftlr-proof-cases.tex
Added a missing todo in FTLR proof
Added missing update of pc in assumptions of FTLR proof.

@@
     \item $\trust = \trusted$
-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
+    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}\updReg{\pcreg}{c}, \src{\stk}, \src{\ms_\stk})$ and
       \begin{itemize}           
@@
       \]
-      Here we need to use the induction hypothesis.
+      Here we need to use the induction hypothesis. \lau{TODO: the reasonability
+      requirements of the FTLR seems to be an issue here.}
     \end{itemize}

### 2018-06-13 fully-abstract-comp/technical-report.tex
Updated semantics of scall, so it matches what the implementation does when one of the registers is the temp register.
typo
Refactored one case of the value relation slightly.
Tentative reformulation of reasonability requirements...

--- 2018-06-12 fully-abstract-comp/technical-report.tex
@@
 
+\subsection{Alternative formulation of reasonability?}
+\label{sec:reasonability}
+
+\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
+  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
+}
+\dominique{TODO: better name than ``reasonable'' for all of this.}
+Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
+We define that a word $w$ is reasonable if the following implications hold.
+\begin{itemize}
+\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ then $[\baddr,\eaddr] \# \ta$
+\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
+\end{itemize}
+
+We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ and data capability $c_{\mathrm{data}}$ behaves reasonably with stack assumption $\ms_\stk'$ up to $n$ steps if the following implication holds:
+
+For any $\Phi$ such that
+  \begin{itemize}
+  \item $\Phi.\reg(\pcreg) = c$
+  \item $\Phi.\reg(\rdata) = c_{\mathrm{data}}$
+  \item $\Phi.\reg(r)$ is reasonable for all $r \not\in \{\pcreg,\rdata\}$
+  \item $\Phi.\mem(a)$ is reasonable for all $a$
+  \item $\Phi.\ms_\stk$ is reasonable for all $a$
+  \item $\Phi.\ms_\stk|_{\dom(\ms_\stk')} = \ms_\stk'$
+  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
+  \end{itemize}
+  We have the following implications:
+  \begin{itemize}
+  \item[Guarantee stack base address before call] If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
+    \end{itemize}
+    Then all of the following hold:
+    \begin{itemize}
+    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
+    \item $r_1' \neq \rtmp{1}$
+    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behave reasonably with stack assumption $\Phi.\ms_\stk|_{[\eaddr_\stk,\aaddr_\stk]}$ up to $n-i$ steps
+    \end{itemize}
+  \item[Use return seals only for calls] If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
+    \end{itemize}
+    Then one of the following holds:
+    \begin{itemize}
+    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
+    \item $\sigma \in \sigcloss$.
+    \end{itemize}
+  \item[Don't store private stuff...] If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
+    \end{itemize}
+    Then $\Phi.\reg(r_2)$ is reasonable
+  \item[Don't leak private stuff...] If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
+    \item $\src{\Phi'}$ is in $\ta$
+    \item $\src{\Phi''}$ is not in $\ta$
+    \end{itemize}
+    Then $\Phi.\reg(r)$ is reasonable for all $r$ (including $\pcreg$)
+  \end{itemize}
+
+We say that a component $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$ is reasonable if the following hold:
+For all $\sealed{\sigma,\vsc},\sealed{\sigma,\vsc'} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$ and $\nonExec{\vsc'}$, we have that $\vsc$ and $\vsc'$ behave reasonably with stack assumption $\emptyset$ up to any number of steps $n$.
+
 \subsection{Fundamental Theorem of Logical Relations}

### 2018-06-14 fully-abstract-comp/technical-report.tex
Reformatted proposal.
Fix typo.
Fixes for Lau's concerns in enw reasonability definition.

@@
     \begin{array}[t]{l}
+      r_1 \neq \rtmp{1} \tand r_2 \neq \rtmp{1} \\
       \Phi(r_1) = \sealed{\sigma_1,c_1} \tand \\
@@
     \begin{array}{l}
+      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
       \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
       \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
-      \quad (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
       \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
@@
 Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
-We define that a word $w$ is reasonable if the following implications hold.
+We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if the following implications hold.
 \begin{itemize}
@@
 \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is readable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$
 \end{itemize}
@@
   \item $\Phi.\reg(\rdata) = c_{\mathrm{data}}$
-  \item $\Phi.\reg(r)$ is reasonable for all $r \not\in \{\pcreg,\rdata\}$
-  \item $\Phi.\mem(a)$ is reasonable for all $a$
-  \item $\Phi.\ms_\stk$ is reasonable for all $a$
+  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg,\rdata\}$
+  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
   \item $\Phi.\ms_\stk|_{\dom(\ms_\stk')} = \ms_\stk'$
@@
   \begin{itemize}
-  \item[Guarantee stack base address before call] If
+  \item \emph{Guarantee stack base address before call} If
     \begin{itemize}
@@
     \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
-    \item $r_1' \neq \rtmp{1}$
+    \item $r_1 \neq \rtmp{1}$
     \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behave reasonably with stack assumption $\Phi.\ms_\stk|_{[\eaddr_\stk,\aaddr_\stk]}$ up to $n-i$ steps
     \end{itemize}
-  \item[Use return seals only for calls] If
+  \item \emph{Use return seals only for calls} If
     \begin{itemize}
@@
     \end{itemize}
-  \item[Don't store private stuff...] If
+  \item \emph{Don't store private stuff...} If
     \begin{itemize}
@@
     \end{itemize}
-    Then $\Phi.\reg(r_2)$ is reasonable
-  \item[Don't leak private stuff...] If
+    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
+  \item \emph{Don't leak private stuff...} If
     \begin{itemize}
@@
     \end{itemize}
-    Then $\Phi.\reg(r)$ is reasonable for all $r$ (including $\pcreg$)
+    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
   \end{itemize}
@@
   \item If $\npair{A} \in \execCond{W}$, then $\npair{A} \in \execCond{}(\purePart{W})$
-  \item If $\npair{(w_1,w_2)} \in \lrv(H_\sigma,W)$ and ($\nonLinear{w_1}$ or
+  \item If $\npair{(w_1,w_2)} \in \lrv(W)$ and ($\nonLinear{w_1}$ or
     $\nonLinear{w_2}$), then
@@
   \[
-    r_1' \neq \rtmp{1}
+    r_1 \neq \rtmp{1}
   \]

### 2018-06-14 fully-abstract-comp/technical-report.tex
Oops, forgot something in the new definition.

@@
 \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is readable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is readable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
 \end{itemize}

### 2018-06-14 fully-abstract-comp/technical-report.tex
Suggestion for simplification of the reasonability condition.

--- 2018-06-14 fully-abstract-comp/technical-report.tex
@@
 \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is readable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$
-\end{itemize}
-
-We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ and data capability $c_{\mathrm{data}}$ behaves reasonably with stack assumption $\ms_\stk'$ up to $n$ steps if the following implication holds:
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
+\end{itemize}
+
+We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps if the following implication holds:
 
@@
   \item $\Phi.\reg(\pcreg) = c$
-  \item $\Phi.\reg(\rdata) = c_{\mathrm{data}}$
-  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg,\rdata\}$
+  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$
   \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
-  \item $\Phi.\ms_\stk|_{\dom(\ms_\stk')} = \ms_\stk'$
   \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
@@
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behave reasonably with stack assumption $\Phi.\ms_\stk|_{[\eaddr_\stk,\aaddr_\stk]}$ up to $n-i$ steps
+    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
     \end{itemize}
@@
 For all $\sealed{\sigma,\vsc},\sealed{\sigma,\vsc'} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$ and $\nonExec{\vsc'}$, we have that $\vsc$ and $\vsc'$ behave reasonably with stack assumption $\emptyset$ up to any number of steps $n$.
+\lau{This need to change with the proposed simplification of the reasonability update.}
 

### 2018-06-14 fully-abstract-comp/technical-report.tex
Sketch of a lemma that could make the call case for the FTLR proof work.
First stab at proving Lau's new lemma.

@@
 % DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
-% Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
+% Where $S\text{ is a separation of $W$}$ means that no two partitionjjjjshare the same ``spatial'' region (whatever that is).
 %\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
@@
 \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ then $[\baddr,\eaddr] \# \ta$
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$, then $[\baddr,\eaddr] \# \dom(\mscode)$
 \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
@@
 \lau{This need to change with the proposed simplification of the reasonability update.}
+
+\begin{lemma}
+  \label{lem:codereg-Hsigma-not-lrv}
+  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma \not\nequal \lrv$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:codereg-not-read-write}
+  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$ and $a \in \dom(\code)$, then
+  \begin{itemize}
+  \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
+  \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}
+  If
+  \begin{itemize}
+  \item $\npair{w,\_} \in \lrv(W_w)$
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
+  \item $W_w \oplus W_M$ is defined
+  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
+  \item $\dom(\code) \subseteq \ta$
+  \item $n > 0$
+  \end{itemize}
+  then, with respect to $\code,\ta,\sigrets,\sigcloss$,
+  $w$ is reasonable up to $n$ steps in memory $\Phi.\mem$.
+\end{lemma}
+\begin{proof}
+  We need to prove four implications:
+  \begin{itemize}
+  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$:
+
+    Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\sigrets \cup \sigcloss)$.
+    Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.
+
+    From $\npair{w,\_} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.
+
+    From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.
+
+    From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
+
+    From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
+
+    From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
+    However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
+
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$, then $[\baddr,\eaddr] \# \ta$:
+
+    Assume that $a \in [\baddr,\eaddr] \cap \ta$ and we will prove a contradiction.
+
+    From $\npair{w,\_} \in \lrv(W_w)$, it follows that $\npair{\{a\}}\in\readCond{\normal,W_w}$ or $\npair{\{a\}}\in\writeCond{\normal,W_w}$.
+
+    This contradicts Lemma~\ref{lem:codereg-not-read-write}.
+
+  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable:
+    TODO
+
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
+    TODO
+
+  \end{itemize}
+\end{proof}
+
+\begin{lemma}
+  If
+  \begin{itemize}
+  \item $\npair{\reg_S,\_} \in \lrr(W_w)$
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
+  \item $W_w \oplus W_M$ is defined
+  \item $\pwheap[W_w](r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
+  \item $\dom(\code) \subseteq \ta$
+  \item $n > 0$
+  \end{itemize}
+  then, with respect to $\code,\ta,\sigrets,\sigcloss$,
+  \begin{itemize}
+  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$
+  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
+  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  We prove the three results separately:
+  \begin{itemize}
+  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$:
+
+    We have that $W = \bigoplus_{r \in \RegName \setminus \{\pcreg\}} S(r)$ and $\npair{(\reg_S(r),\reg_T(r))} \in \lrv(S(r))$ for all $r \not\in \{\pcreg\}$.
+
+    TODO
+
+  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
+    TODO
+  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
+    TODO
+  \end{itemize}
+\end{proof}
 

### 2018-06-15 fully-abstract-comp/ftlr-proof-cases.tex
Implement the alternative FTLR proof structure that I suggested

--- 2018-06-12 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \item If $\npair{A} \in \xReadCond{W}$, then $\npair[n']{A} \in \xReadCond{W}$.
-  \item If $\npair{(w,w')} \in \lrvg{\trust}{W}$, then $\npair[n']{(w,w')} \in \lrvg{\trust}{W}$.
+  \item If $\npair{(w,w')} \in \lrvg{\trust}(W)$, then $\npair[n']{(w,w')} \in \lrvg{\trust}(W)$.
   \item If $\npair{(w,w')} \in \lrrg{\trust}(W)$, then $\npair[n']{(w,w')} \in \lrrg{\trust}(W)$.
-  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}(W)$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}(W)$.
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
   \end{itemize}
@@
   \item $\xReadCond{W} \nequal \xReadCond{W'}$.
+  \item If $\stackReadCond{\lin,W} \nequal \stackReadCond{\lin,W'}$.
+  \item If $\writeCond{\lin,W} \nequal \writeCond{\lin,W'}$.
+  \item If $\stackWriteCond{\lin,W} \nequal \stackWriteCond{\lin,W'}$
+  \item If $\execCond{\lin,W} \nequal \execCond{\lin,W'}$.
+  \item If $\xReadCond{W} \nequal \xReadCond{W'}$.
+  \item If $\lrvg{\trust}(W) \nequal \lrvg{\trust}(W')$.
+  \item If $\lrrg{\trust}(W) \nequal \lrrg{\trust}(W')$.
+  \item For all $n' \le n$, $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$ iff $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W'}$.
   \item ...
@@
 \subsection{FTLR proof}
-By induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
+
+\begin{lemma}
+  \label{lem:ftlr-internal-lemma}
+  If
+  \begin{itemize}
+  \item One of the following sets of requirements holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \end{itemize}
+  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$
+  \item $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$
+  \item $\memSat{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$
+  \item Theorem~\ref{thm:ftlr} holds for all $n' < n$.
+  \end{itemize}
+  Then
+  \[
+    \npair{(\Phi_S,\Phi_T)} \in \lro
+  \]
+\end{lemma}
 \begin{proof}
-  Assume
-  \begin{itemize}
-  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
-  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
-  \end{itemize}
-  and one of the following sets of requirements holds:
-  \begin{enumerate}[label={\roman*)}]
-  \item \begin{itemize}
-    \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \trusted$
-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}, \src{\stk}, \src{\ms_\stk})$ and
-      \begin{itemize}           
-      \item $\src{\Phi}$ checks $\stkb$ before calls in $\ta$ (Definition~\ref{def:check-stack-addr-before-call}).
-      \item $\src{\Phi}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ (Definition~\ref{def:use-return-seals-call}).
-      \item $\src{\Phi}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ (Definition~\ref{def:handle-trusted-seals-properly}).
-      \item $\src{\Phi}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ (Definition~\ref{def:never-store-seal-code-cap}).
-      \end{itemize}
-    \end{itemize}
-  \item
-    \begin{itemize}
-    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
-    \item $\trust = \untrusted$
-    \end{itemize}
-  \end{enumerate}
-and show
-\[
-\npair{(c,c)} \in \lreg{\trust}(W)
-\]
-For $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
-
-Let $n' \leq n$ be given and assume
-\begin{enumerate}
-\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$ \label{item:reg-ass}
-\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
-\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
-\end{enumerate}
-Further let
-\begin{itemize}
-\item $\Phi_S = (\ms_S,\reg_S\update{\pcreg}{c},\stk,\ms_\stk)$
-\item $\Phi_T = (\ms_T,\reg_T\update{\pcreg}{c})$
-\end{itemize}
-and show
-\[
-  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
-\]
+  TODO: update...
+
+  By general induction on $n$, i.e. we can assume that the property already holds for $n' < n$.
+
   In order to do this, we first, we prove that one of the following holds:
@@
   \\\\
-In case \ref{case:ftlr:get-and-arith-op}, we use the induction hypothesis to get
-\[
-  \npair[n'-1]{(((\perm,\lin),\baddr,\eaddr,\aaddr+1),((\perm,\lin),\baddr,\eaddr,\aaddr+1))} \in \lre(W)
-\]
-This is possible by using the initial assumptions along with Lemma \ref{lem:downwards-closed}.
-By Lemma~\ref{lem:downwards-closed}, we know $\memSat[n'-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
-By Lemma~\ref{lem:downwards-closed} and the fact that integers are always safe,
-we know $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$.
-This allows us to conclude
+  In case \ref{case:ftlr:get-and-arith-op}, we use the induction hypothesis to conclude
   \[
@@
   \]
-as desired.
+  from the following facts:
+  \begin{itemize}
+  \item One of the following sets of requirements holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \end{itemize}:
+    Follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep[1] \Phi_S'$ with Lemma~\ref{lem:ec-reasonable-antired}.
+    \dominique{not actually true: reasonability definitions need to be made downward closed using quantification over smaller $n'$?}
+  \item $\trust = \trusted \vee \npair[n'-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$:
+    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}.
+  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$:
+    Follows from the definition of $\updPcAddr{}$ and the assumptions of this case.
+  \item $\npair[n'-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}.
+  \item $\npair[n' - 1]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
+    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed} and the fact that integers are always in $\lrvg{\trust}$ by definition.
+  \item $\memSat[n'-1]{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
+    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}.
+  \item Theorem~\ref{thm:ftlr} holds for all $n'' < n'-1$:
+    follows from the corresponding assumption of this lemma, since $n' - 1 < n$.
+  \end{itemize}
   \\\\
@@
         \]
-      Here we need to use the induction hypothesis.
+        Here we need to use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$.
+        TODO: continue
+        \lau{TODO: the reasonability requirements of the FTLR seems to be an issue here.}
+        \dominique{should now be fine?}
       \end{itemize}
@@
   Case: \ref{case:ftlr:scall:mem}, \lau{TODO}
-
+\end{proof}
+
+\begin{proof}[Proof of Theorem~\ref{thm:ftlr}]
+  By general induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
+  Assume
+  \begin{itemize}
+  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
+  \end{itemize}
+  and one of the following sets of requirements holds:
+  \begin{enumerate}[label={\roman*)}]
+  \item \begin{itemize}
+    \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \trusted$
+    \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
+    \end{itemize}
+  \item
+    \begin{itemize}
+    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
+    \item $\trust = \untrusted$
+    \end{itemize}
+  \end{enumerate}
+and show
+\[
+\npair{(c,c)} \in \lreg{\trust}(W)
+\]
+For $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
+
+Let $n' \leq n$ be given and assume
+\begin{enumerate}
+\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$ \label{item:reg-ass}
+\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
+\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
+\end{enumerate}
+Further let
+\begin{itemize}
+\item $\Phi_S = (\ms_S,\reg_S\update{\pcreg}{c},\stk,\ms_\stk)$
+\item $\Phi_T = (\ms_T,\reg_T\update{\pcreg}{c})$
+\end{itemize}
+and show
+\[
+  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
+\]
+
+By Lemma~\ref{lem:ftlr-internal-lemma}, it suffices to show that:
+\begin{itemize}
+\item $\Phi_S$ is reasonable up to $n'$ steps:
+  TODO
+\item $\trust = \trusted \vee \npair[n']{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$:
+  TODO
+\item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr)$:
+  TODO
+\item $\npair[n']{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+  TODO
+\item $\npair[n']{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\tst}(W_R)$:
+  TODO
+\item $\memSat[n']{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
+  TODO
+\item Theorem~\ref{thm:ftlr} holds for all $n'' < n'$:
+  Follows from our induction hypothesis since $n' \le n$.
+\end{itemize}
 \end{proof}

### 2018-06-15 fully-abstract-comp/technical-report.tex
typo
Added requirement in V-relation for sealed capabilities that H sigma must be the one from code region.
Consider whether the requirement should be that it is the code region.
Implement the alternative FTLR proof structure that I suggested

@@
 
-We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps if the following implication holds:
-
-For any $\Phi$ such that
+We say that an execution configuration $\Phi$ is reasonable up to $n$ steps if:
+  \begin{itemize}
+  \item \emph{Guarantee stack base address before call} If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
+    \item $i \le n$
+    \end{itemize}
+    Then all of the following hold:
+    \begin{itemize}
+    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
+    \item $r_1 \neq \rtmp{1}$
+    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
+    \end{itemize}
+  \item \emph{Use return seals only for calls} If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
+    \item $i \le n$
+    \end{itemize}
+    Then one of the following holds:
+    \begin{itemize}
+    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
+    \item $\sigma \in \sigcloss$.
+    \end{itemize}
+  \item \emph{Don't store private stuff...} If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
+    \item $i \le n$
+    \end{itemize}
+    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
+  \item \emph{Don't leak private stuff...} If
+    \begin{itemize}
+    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
+    \item $\src{\Phi'}$ is in $\ta$
+    \item $\src{\Phi''}$ is not in $\ta$
+    \item $i \le n$
+    \end{itemize}
+    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
+  \end{itemize}
+
+  \begin{lemma}
+    \label{lem:ec-reasonable-antired}
+    If
+    \begin{itemize}
+    \item $\Phi$ is reasonable up to $n$ steps
+    \item $\Phi \nstep[i]{\gc} \Phi'$
+    \end{itemize}
+    Then
+    \begin{itemize}
+    \item $\Phi'$ is reasonable up to $n-i$ steps
+    \end{itemize}
+  \end{lemma}
+  \begin{proof}
+    Follows easily by definition.
+  \end{proof}
+
+We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps if for any $\Phi$ such that
   \begin{itemize}
@@
   \end{itemize}
-  We have the following implications:
-  \begin{itemize}
-  \item \emph{Guarantee stack base address before call} If
-    \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
-    \end{itemize}
-    Then all of the following hold:
-    \begin{itemize}
-    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
-    \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
-    \end{itemize}
-  \item \emph{Use return seals only for calls} If
-    \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
-    \end{itemize}
-    Then one of the following holds:
-    \begin{itemize}
-    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
-    \item $\sigma \in \sigcloss$.
-    \end{itemize}
-  \item \emph{Don't store private stuff...} If
-    \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
-    \end{itemize}
-    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
-  \item \emph{Don't leak private stuff...} If
-    \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
-    \item $\src{\Phi'}$ is in $\ta$
-    \item $\src{\Phi''}$ is not in $\ta$
-    \end{itemize}
-    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
-  \end{itemize}
+  We have that $\Phi$ is reasonable.
 
 We say that a component $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$ is reasonable if the following hold:
-For all $\sealed{\sigma,\vsc},\sealed{\sigma,\vsc'} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$ and $\nonExec{\vsc'}$, we have that $\vsc$ and $\vsc'$ behave reasonably with stack assumption $\emptyset$ up to any number of steps $n$.
+For all $\sealed{\sigma,\vsc} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$, we have that $\vsc$ behaves reasonably up to any number of steps $n$.
 \lau{This need to change with the proposed simplification of the reasonability update.}
@@
     \item $\trust = \trusted$
-    \item For all $\npair{\stpair{\reg}{\reg}} \in \lrrtrusted(W)$ and $\memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W}$ and $\ms_f$, we have $\src{\Phi} = (\src{\ms_S} \uplus \ms_f, \src{reg_S}\updReg{\pcreg}{((\rx,\lin),\baddr,\eaddr,\aaddr)}, \src{\stk}, \src{\ms_\stk})$ and
-      \begin{itemize}
-      \item $\src{\Phi}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-      \item $\src{\Phi}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-      \item $\src{\Phi}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-      \item $\src{\Phi}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
-      \end{itemize}
+    \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
     \end{itemize}

### 2018-06-15 fully-abstract-comp/ftlr-proof-cases.tex
Remove trailing whitespace
Fix latex compilation.

@@
     \end{itemize}:
-    Follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep[1] \Phi_S'$ with Lemma~\ref{lem:ec-reasonable-antired}.
+    Follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep{1} \Phi_S'$ with Lemma~\ref{lem:ec-reasonable-antired}.
     \dominique{not actually true: reasonability definitions need to be made downward closed using quantification over smaller $n'$?}
@@
   \end{itemize}
-  \\\\
+
   \lau{TODO: look through this case and find the place where the ''no seal cross boundary thing'' should have been used.}
@@
   TODO
-\item $\npair[n']{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\tst}(W_R)$:
+\item $\npair[n']{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
   TODO

### 2018-06-15 fully-abstract-comp/ftlr-proof-cases.tex
Minor fix.
merge?

@@
   TODO: update...
-
-  By general induction on $n$, i.e. we can assume that the property already holds for $n' < n$.
-
-  In order to do this, we first, we prove that one of the following holds:
+  By general induction on $n$, i.e. we can assume that the lemma already holds for $n' < n$.
+
+  In order to prove this, we first, we prove that one of the following holds:
   \lau{TODO: add $\neq$ failed to the cases that are not failed.}
@@
     \item $\Phi_T \step \Phi_T' \neq \failed$
-    \item $\Phi_T' = \updPc{\Phi_T\updReg{\rtmp{1}}{z}}$
+      % domi: this condition doesn't make syntactic sense and is also not used.
+    % \item $\Phi_T' = \updPc{\Phi_T\updReg{\rtmp{1}}{z}}$
     \item $z \in \ints$
@@
     \item $\Phi_S''.\ms_\stk = \Phi_S.\ms_\stk -
-      \Phi_S.\ms_\stk|_{\aaddr_\stk,\eaddr_\stk}$
+      \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}$
     \item $\baddr \leq \aaddr+\offpc \leq \eaddr$
@@
       \end{itemize}
-
-    \end{itemize}
+    \end{itemize}
+
+\end{description}
+This concludes case \ref{case:ftlr:scall:reg}
+
+Case: \ref{case:ftlr:scall:mem}, we need to show
+
+\begin{equation}
+  \memSat[n'-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}\label{eq:ftlr:call:memsat-case}
+\end{equation}
+We have the following memory satisfaction on the initial memory:
+\begin{equation}
+  \memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}\label{eq:ftlr-call-og-mem-sat}
+\end{equation}
+By definition of memory satisfaction, this gives us:
+\begin{itemize}
+\item $\stk = (\opc_1,\ms_1) :: \dots :: (\opc_m,\ms_m)$
+\item $\ms_1 \# \dots \# \ms_m \# \ms_S \# \ms_\stk$
+\item $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$
+\item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
+  \begin{itemize}
+  \item $\ms_S = \ms_{f,S} \uplus \ms_S'$
+  \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
+  \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
+  \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})$
+  \end{itemize}
+\end{itemize}
+We need to consider the memories of $\Phi_S''$ and $\Phi_T''$.
+The memories of the initial configurations have only changed little:
+\[
+  \Phi_T''.\mem = \Phi_T''.\mem\update{\aaddr_\stk}{42}
+\]
+and 
+\[
+\Phi_S''.\ms_\stk = \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
+\]
+and
+\[
+  \Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk
+\]
+with $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
+We need to show the memory satisfaction in \ref{eq:ftlr:call:memsat-case}. Say
+$(\opc_0,\ms_0) = ((\aaddr+\calllen),\ms_{\stk\_\priv,S})$. In this case we have
+to argue 
+\[
+  \ms_{\stk\_\priv,S} \# \ms_1 \# \dots \# \ms_m \# \ms_S \# \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
+\]
+which follows from Assumption~\ref{eq:ftlr:call:memsat-case} and the fact that the memory we remove from $\Phi_S.\ms_\stk$ is exactly what we add to the private stack.
+\lau{This is a bit more involved than I make it to be here, but the fact that we also have a safety assumption on the register-file with the stack pointer means that all of the memory removed from the free stack is actually there.}
 

### 2018-06-15 fully-abstract-comp/technical-report.tex
merge?
Changed the comp code definition so all the seals in the judgement must be used (but they can be used multiple times for seals, but they still need to be unique per call).
A few of the suggested changes.

@@
   \inference{
-    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} & [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)
+    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} & [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
   }{
@@
     \sigrets \mathrel{\#} \sigcloss \\
-    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
-    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
+    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal}^3 \ldotp
+    \sigrets = \biguplus_{a \in \dom(\mscode)} \phi_2(d_\sigma(a)) \tand\\
+    \sigrets = \bigcup_{a \in \dom(\mscode)} \phi_1(d_\sigma(a)) \tand\\
+    \sigcloss = \bigcup_{a \in \dom(\mscode)} \phi_3(d_\sigma(a)) \tand\\
     \forall a \in \dom(\mscode) \ldotp 
-    \sigrets,d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+    \pi_1(d_\sigma(a)),\pi_2(d_\sigma(a)),\pi_3(d_\sigma(a)) \vdash_{\mathrm{comp-code}} \mscode,a
   }{
@@
 % DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
-% Where $S\text{ is a separation of $W$}$ means that no two partitionjjjjshare the same ``spatial'' region (whatever that is).
+% Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
 %\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
@@
       (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
-      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \quad H_\sigma = H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \tand \\
       \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
@@
     \begin{array}{l}
-      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) =_n \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
+      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
       \hfill\dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)
@@
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-    (\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
+      (\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee \\
+      \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
@@
 \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$, then $[\baddr,\eaddr] \# \dom(\mscode)$
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\sigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
 \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
@@
   \label{lem:codereg-not-read-write}
-  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$ and $a \in \dom(\code)$, then
+  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and $a \in \dom(\code)$ and $\memSat{\code \uplus \ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\sigrets \neq \emptyset$, then
   \begin{itemize}
   \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
-  \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
+    % \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
+    % It seems unnecessary to show the part about the write condition as all non-no-perm permissions have read.
   \end{itemize}
@@
   \begin{itemize}
-  \item $\npair{w,\_} \in \lrv(W_w)$
+  \item $\npair{(w,\_)} \in \lrv(W_w)$
   \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
@@
 
-    From $\npair{w,\_} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.
+    From $\npair{(w,\_)} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.
 
@@
 
-    From $\npair{w,\_} \in \lrv(W_w)$, it follows that $\npair{\{a\}}\in\readCond{\normal,W_w}$ or $\npair{\{a\}}\in\writeCond{\normal,W_w}$.
+    From $\npair{(w,\_)} \in \lrv(W_w)$, it follows that $\npair{\{a\}}\in\readCond{\normal,W_w}$.
 
@@
   \begin{itemize}
-  \item $\npair{\reg_S,\_} \in \lrr(W_w)$
+  \item $\npair{(\reg_S,\_)} \in \lrr(W_w)$
   \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$

### 2018-06-16 fully-abstract-comp/technical-report.tex
Changed and proved one of the lemmas needed to prove the safe implies reasonable lemma.

@@
   \label{lem:codereg-Hsigma-not-lrv}
-  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma \not\nequal \lrv$.
+  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma \sigma \not\nequal \lrv$.
 \end{lemma}
@@
   \label{lem:codereg-not-read-write}
-  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and $a \in \dom(\code)$ and $\memSat{\code \uplus \ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\sigrets \neq \emptyset$, then
+  If $n > 0$, $\dom(\code) \subseteq \ta$, $\gc = (\ta,\stkb)$, $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and $a \in \dom(\code)$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\sigrets \neq \emptyset$, then
   \begin{itemize}
@@
   \end{itemize}
+\end{lemma}
+\begin{proof}
+  For the sake of contradiction assume $\npair{\{a\}} \in \readCond{\normal,W}$.
+  Then there exists $S \subseteq \addressable{\lin, \pwheap}$ and $R : S \fun
+  \powerset{\nats}$ such that
+  \begin{itemize}
+  \item $\biguplus_{r\in S} R(r) \supseteq A$
+  \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
+  \end{itemize}
+  We claim that $r_\code \in S$\lau{TODO: show this claim}.
+  If this is indeed the case, then we have $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$ (we must have $R(r_\code) = \dom(\code)$ as the two regions couldn't be $\nequal$ otherwise).
+  This means that for all $a \in \dom(\code)$ and $n' < n$ it must be the case that $\npair[n']{(\code(a),\code(a))} \in \lrv(W_a)$ for all $W_a$ that are compatible w.r.t.
+  $\oplus$ and combine to be $W$.
+  We know that $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ from which we can deduce $\npair{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \xi^{-1}(W_\code)$ which in particular gives us $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code$.
+  This in turns entails that for $\sigma \in \sigrets$ there exists some $a' \in \dom(\code)$ such that $\code(a') = \seal{\sigma_\aaddr,\sigma_\baddr,\sigma_\eaddr}$ and $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
+  This means that for such a $\sigma$ and $a'$, it must be that case that
+  $\npair[n']{(\seal{\sigma_\aaddr,\sigma_\baddr,\sigma_\eaddr},\seal{\sigma_\aaddr,\sigma_\baddr,\sigma_\eaddr})}
+  \in \lrv(W_{a'})$ which gives us
+  $\codereg{\sigrets,\sigcloss,\code,\gc}.H_\sigma  \; \sigma \nequal \lrv$
+  contradicting Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
+
+  It remains to show $r_\code \in S$.
+  To show this, we assume for contradiction that $r_\code \not\in S$.
+  This means that there exists $r \neq r_\code$ such that $a \in R(r)$ and $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$.
+  Because we know $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, it must be possible to split up $\ms_S$ in disjoint memory segments that each (paired with a target memory segment) satisfies exactly one region.
+  In this case, both $r$ and $r_\code$ needs a memory segment that has $a$ in the domain which means that no such disjoint union can exist which gives us a contradiction.
+\end{proof}
+
+\begin{lemma} 
+  \label{lem:code-reg-and-mem-sat}
+  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and
+  $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, then there exists $W''$
+  and $W'$ such that $W = W' \oplus W''$ and
+  $\npair{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc \; \xi^{-1}(W')$
 \end{lemma}

### 2018-06-18 fully-abstract-comp/technical-report.tex
Removed a TODO that was resolved.
Technical fix: exact equality breaks non-expansiveness, use n-equality instead.
Some work on the sealed case - it still does not work, but it gives an idea about where the proof fails.

@@
   \end{itemize}
-  We claim that $r_\code \in S$\lau{TODO: show this claim}.
+  We claim that $r_\code \in S$.
   If this is indeed the case, then we have $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$ (we must have $R(r_\code) = \dom(\code)$ as the two regions couldn't be $\nequal$ otherwise).
@@
 
-  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable:
+  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable: By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}$.
+
+    It also follows that $\sigrets \uplus \sigcloss \# \sigrets' \uplus \sigcloss'$.
+
+    This gives us three cases for $\vsc$.
+    The two first cases $\vsc = \retptrd(\_,\_)$ and $\vsc = \retptrc(\_,\_,\_)$ are easily discharged as the reasonability definition puts no requirements on return pointers.
+    For the final case, $\sigma \in \sigcloss$ and either
+    \begin{itemize}
+    \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n']{(\vsc,\vsc')} \in \lrv \;
+      \xi(\hat{W})$; or
+    \item $\dom(\code') \subseteq \ta \tand \npair[n']{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})$
+    \end{itemize}
+    \lau{At this point, it seems like should have done the proof by induction,
+      but this would not be possible with the $n>0$ assumption (at least not if
+      we want to use the IH). In the latter case, it is probably difficult to
+      show that the something is in the trusted value-relation. If we assume
+      $\ta = \dom(\code)$, then this may be discharged by contradiction (we may
+      somewhere need to assume that $\ta \neq \emptyset$).}
     TODO

### 2018-06-18 fully-abstract-comp/technical-report.tex
Changed reasonability to always hold for n=0.

@@
       \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad H_\sigma = H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \tand \\
+      \quad H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \tand \\
       \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
@@
 Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
-We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if the following implications hold.
+We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if $n=0$
+or the following implications hold.
 \begin{itemize}
@@
 \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\sigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
-\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
+\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
 \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
@@
     \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
-    \item $i \le n$
+    \item $0 < i \le n$
     \end{itemize}
@@
     \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
-    \item $i \le n$
+    \item $0 < i \le n$
     \end{itemize}
@@
     \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
-    \item $i \le n$
+    \item $0 < i \le n$
     \end{itemize}
@@
     \item $\src{\Phi''}$ is not in $\ta$
-    \item $i \le n$
+    \item $0 < i \le n$
     \end{itemize}

### 2018-06-18 fully-abstract-comp/technical-report.tex
first stap at global ret and clos seals.

@@
 \[
-  \lrvtrusted[\square,(\ta,\stkb)](W) =
+  \lrvtrusted[\square,(\ta,\stkb,\gsigrets,\gsigcloss)](W) =
   \begin{array}[t]{l}
@@
     \begin{array}{l}
-      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
-      \hfill\dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)
+      \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
+      \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
+      \quad \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss
     \end{array}
@@
       \perm \sqsubseteq \rx \wedge \\
-      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,(\ta,\stkb)]{W}
+      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
+      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
     \end{array}
@@
 \begin{equation*}
-  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
     \begin{array}{l}
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-      (\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee \\
+      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \trust = \trusted) \vee \\
       \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
@@
       \begin{array}{l}
+        \sigrets \subseteq \gsigrets \tand \\
         \dom(\code) \subseteq \ta \tand\\
@@
         (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
+        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
       \end{array}
@@
 \dominique{TODO: better name than ``reasonable'' for all of this.}
-Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
+Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
 We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if $n=0$
@@
 \begin{itemize}
-\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\sigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
-\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
+\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
+\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
 \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
@@
     \begin{itemize}
-    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
-    \item $\sigma \in \sigcloss$.
+    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$ 
+    \item $\sigma \in \gsigcloss$.
     \end{itemize}

### 2018-06-18 fully-abstract-comp/ftlr-proof-cases.tex
compilation error
Remove the non-sensical requirement everywhere.
Add missing assumption to new lemma.

@@
     In the first case, $\vsc_S'$ is a $\retptrc$ and $\vsc_T'$ is a capability with permission $\rx$.
-    In this case, $\xjumpResult{}{}{}$ will give to failed configurations which is trivially in the observation relation.
+    In this case, $\xjumpResult{}{}{}$ will produce failed configurations which are trivially in the observation relation.
     In the next case, it is required that $\sigma' \in \sigcloss'$, but this cannot be the case as \lau{TODO I think this case should follow from $H^\var{code}$ and the comp-code judgement, but it does not.
-      Specifically, the comp-code judgement (for now) does not require $\sigrets'$ and $\sigcloss'$ to be disjointed.}
+      Specifically, the comp-code judgement (for now) does not require $\sigrets'$ and $\sigcloss'$ to be disjoint.}
 
@@
     By the above, the fact that the code capability pair is in $H_\sigma^\var{code}$ (\ref{eq:code-caps-in-h-sigma}) and Lemma~\ref{lem:unique-ret-seals}, we get $\aaddr' = \aaddr$.
-    This means that $W_M''$ and $W_R''$ has this region.
+    This means that $W_M''$ and $W_R''$ have this region.
 
@@
 
-  \end{description}
   This concludes case \ref{case:ftlr:scall:reg}
@@
   \item \begin{itemize}
-    \item $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $[\baddr,\eaddr] \subseteq \ta$
     \item $\trust = \trusted$
@@
   \item $\mscode([\aaddr'..\aaddr'+\calllen-1]) = \overline{\scall{\offpc',\offsigma'}{r_1'}{r_2'}}$
-  \item $\mscode(\aaddr + \offpc) = \seal{\sigma_b,\sigma_e,\sigma_b}$ and $\sigma = \sigma_b + \offpc$
-  \item $\mscode(\aaddr' + \offpc') = \seal{\sigma_b',\sigma_e',\sigma_b'}$ and $\sigma = \sigma_b' + \offpc'$
+  \item $\mscode(\aaddr + \offpc) = \seal{\sigma_b,\sigma_e,\sigma_b}$ and $\sigma = \sigma_b + \offsigma$
+  \item $\mscode(\aaddr' + \offpc') = \seal{\sigma_b',\sigma_e',\sigma_b'}$ and $\sigma = \sigma_b' + \offsigma'$
+  \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$
   \end{itemize}

### 2018-06-18 fully-abstract-comp/technical-report.tex
Fix what seems to be an omission in Lau's most recent commit.
Remove non-sensical requirement in FTLR.
Finish a few minor proofs.

@@
     \begin{array}{l}
+      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
       \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal \lrv
+      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
     \end{array}
@@
         (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
+        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
       \end{array}
@@
 \begin{lemma}
-  \label{lem:codereg-Hsigma-not-lrv}
-  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma \sigma \not\nequal \lrv$.
+  \label{lem:lrvtrusted-not-nequal-lrvuntrusted}
+  If $n > 0$
+  then
+  $\lrv \not\nequal \lrvg{\trusted}$
 \end{lemma}
 \begin{proof}
-  TODO
+  Take $\sigma \in (\gsigrets\cup \gsigcloss)$, $w = \seal{\sigma,\sigma,\sigma}$, $W$ such that $W(r) = \codereg{\sigrets,\sigcloss,\code,\gc}$, $\sigrets \subseteq \gsigrets$, $\sigcloss\subseteq \gsigcloss$, $\dom(\code) \subseteq \ta$ and $W$ empty otherwise.
+  Then we have that $\npair[0]{(w,w)}\in \lrvg{\trusted}(W)$ but not $\npair[0]{(w,w)}\in\lrv(W)$.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:codereg-Hsigma-not-lrv}
+  If $\dom(\code) \subseteq \ta$, $\sigrets \subseteq \gsigrets$, $\sigcloss \subseteq \gsigcloss$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)$.
+\end{lemma}
+\begin{proof}
+  Consider the following two cases:
+  \begin{itemize}
+  \item $\sigma \in \sigrets$:
+
+    In this case, we have that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but we can show that it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$ by taking a $\hat{W}$ such that the following hold:
+    \begin{itemize}
+    \item $\sigrets \subseteq \gsigrets$: assumption.
+    \item $\dom(\code) \subseteq \ta$: assumption.
+    \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
+      It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions. 
+    \end{itemize}
+
+  \item $\sigma \in \sigcloss$: This follows easily from the definition of $\codereg{}$ and Lemma~\ref{lem:lrvtrusted-not-nequal-lrvuntrusted}.
+  \end{itemize}
 \end{proof}
@@
   and $W'$ such that $W = W' \oplus W''$ and
-  $\npair{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc \; \xi^{-1}(W')$
+  $\npair[n']{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc \; \xi^{-1}(W')$ for all $n' < n$
 \end{lemma}
 \begin{proof}
-  TODO
+  By definition of $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we get that $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
+  $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
+  \lrheap(\pwheap)(W_{\var{heap}})$ for some $\ms_S' \subseteq \ms_S$ and $\ms_{T,\var{heap}}\subseteq \ms_T$.
+
+  By definition of $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
+  \lrheap(\pwheap)(W_{\var{heap}})$, we get an 
+  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $\npair{R_\ms(r_{\var{code}})} \in  \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$.
+
+  Since $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$, this implies that also $\npair[n']{R_\ms(r_{\var{code}})} \in  \codereg{\sigrets,\sigcloss,\code,\gc}.H~ \xi^{-1}(R_W(r_{\var{code}}))$, i.e.
+$\npair[n']{R_\ms(r_{\var{code}})} \in  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc~ \xi^{-1}(R_W(r_{\var{code}}))$ as required.
 \end{proof}

### 2018-06-18 fully-abstract-comp/technical-report.tex
Remove obsolete TODOs.

@@
 \lau{TODOs from last day in Leuven}
-\lau{RESOLVED: components have optional main (pc/data pair)}
-\lau{RESOLVED: linking component and context requires only one main (i.e., it is undefined if both or none of the two expose main.)}
-\lau{RESOLVED?: Component logical relation. Rough sketch is in white-board-notes/leuven-23-3-2018-3.jpg (top left).}
-\dominique{RESOLVED: Compatibility lemma for linking: if $c_1$ and $c_1'$ are related and $c_2$ and $c_2'$, then so should be $c_1\bowtie c_2$ and $c_1' \bowtie c_2'$ (if defined).}
-\lau{RESOLVED: Version of FTLR for programs.}
-\lau{RESOLVED: Compatibility lemma for initial executable configuration construct, i.e.
-a lemma that says ``if two programs are related, then their initial executable configurations will be related''.}
-\lau{RESOLVED: A logical relation for executable configurations.}
-\lau{RESOLVED: contextual equivalence should be defined in terms of initial executable configuration.}
 \lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}

### 2018-06-18 fully-abstract-comp/technical-report.tex
Work on third and fourth case of saftey implies reasonable lemma.

--- 2018-06-18 fully-abstract-comp/technical-report.tex
@@
   \label{lem:codereg-Hsigma-not-lrv}
-  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma \sigma \not\nequal \lrv$.
+  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma \sigma \not\nequal \lrv$.
 \end{lemma}
@@
   \begin{itemize}
+  \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
   \item $\npair{(w,\_)} \in \lrv(W_w)$
@@
   \item $W_w \oplus W_M$ is defined
-  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
+  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
   \item $\dom(\code) \subseteq \ta$
-  \item $n > 0$
-  \end{itemize}
-  then, with respect to $\code,\ta,\sigrets,\sigcloss$,
+  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
+  \end{itemize}
+  then, with respect to $\code,\ta,\gsigrets,\gsigcloss$,
   $w$ is reasonable up to $n$ steps in memory $\Phi.\mem$.
@@
 \begin{proof}
-  We need to prove four implications:
-  \begin{itemize}
-  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$:
-
-    Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\sigrets \cup \sigcloss)$.
+  Induction over $n$.
+  For $n=0$ all words are reasonable.
+  For $n>0$, we need to prove four implications:
+  \begin{itemize}
+  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$:
+
+    Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\gsigrets \cup \gsigcloss)$.
     Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.
@@
 
-  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable: By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}$.
+  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable for $n-1$ steps: By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}$, so in particular for $n-1$.
 
@@
     \begin{itemize}
-    \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n']{(\vsc,\vsc')} \in \lrv \;
-      \xi(\hat{W})$; or
-    \item $\dom(\code') \subseteq \ta \tand \npair[n']{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})$
-    \end{itemize}
-    \lau{At this point, it seems like should have done the proof by induction,
-      but this would not be possible with the $n>0$ assumption (at least not if
-      we want to use the IH). In the latter case, it is probably difficult to
-      show that the something is in the trusted value-relation. If we assume
-      $\ta = \dom(\code)$, then this may be discharged by contradiction (we may
-      somewhere need to assume that $\ta \neq \emptyset$).}
+    \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n-1]{(\vsc,\_)} \in \lrv (W)$; or
+    \item $\dom(\code') \subseteq \ta \tand \sigcloss' \subseteq \gsigcloss \npair{(\vsc,\_)} \in \lrvtrusted(W)$
+    \end{itemize}
+    and in either case $\sigma \in \sigcloss'$
+    In the first case, the result follows from the induction hypothesis.
+    In the second case, we have a contradiction with $\sigma \not\in (\gsigrets\cup \gsigcloss)$.
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
+
+    From $\npair{(w,\_)} \in \lrv(W_w)$, we get $\npair[n]{[b,e]} \in \readCond{\perm,\lin}$.
+
+    From $\npair[n]{[b,e]} \in \readCond{\perm,\lin}$ we get $S\subseteq
+    \addressable{\lin,W}$ and $R : S \fun \powerset{\nats}$ such that
+    \begin{itemize}
+    \item $\biguplus_{r\in S} R(r) \supseteq [b,e]$
+    \item $(\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1)$
+    \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
+    \end{itemize}
+    given $a \in [b,e]$ we know by the above that there exists $r$ such that $a
+    \in R(r)$.
+
+    By $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, we get $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that $\ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r))$ and $\ms_S = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r))$.
+
+    Further, we get $R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $\forall r \in \dom(\activeReg{\pwheap}) \ldotp \npair{R_\ms(r)} \in \pwheap(r).H \; \xi^{-1}(R_W(r))$.
+    For $W_M = W' \oplus W''$ for some $W''$.
+
+    In particular, we have $\npair{R_\ms(r)} \in \pwheap(r).H \; \xi^{-1}(R_W(r))$.
+
+    We know $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$, so $\npair[n']{R_\ms(r)} \in \stdreg{R(r),\gc}{\pur}.H \; \xi^{-1}(R_W(r))$ for $n' < n$.
+    This means that for $n'' < n'$, we have $\npair[n'']{\pi_1(R_\ms(r))(a)} \in \lrv(R_W'(a))$ where $R_W' : [b,e] \fun \World$ such that $\biguplus_{a \in \dom(\pi_1(R_\ms(r)))} R_W'(a) = R_W(r)$.
+
+    \lau{At this point we would like to use the IH, but we do not have $R_W'(a) \uplus W_M$ defined as $R_W'(a)$ is a split of $W_M$.}
+
     TODO
-
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
-    TODO
-
   \end{itemize}
@@
   \item $\dom(\code) \subseteq \ta$
-  \item $n > 0$
-  \end{itemize}
-  then, with respect to $\code,\ta,\sigrets,\sigcloss$,
+  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
+  \end{itemize}
+  then, with respect to $\code,\ta,\gsigrets,\gsigcloss$,
   \begin{itemize}
@@
   \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
-    TODO
+    This follows from $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$.
   \end{itemize}

### 2018-06-18 fully-abstract-comp/ftlr-proof-cases.tex

@@
 \end{lemma}
+\begin{proof}
+  Special case of Lemma~\ref{lem:lro-anti-red-gen}.
+\end{proof}
 
@@
 \end{lemma}
+\begin{proof}
+  Our two languages are deterministic\dominique{Lemma needed for this, I guess..}, so we have that $\Phi_S \sterm{\ta,\stkb}$ iff $\Phi_S \sterm{\ta,\stkb}$ and $\Phi_T \term$ iff $\Phi_T' \term$.
+  It is also easy to show that if $\Phi_S \sterm[i + \min(j_S,j_T)]$, then $\Phi_S' \sterm[i]$ and if $\Phi_T \term[i+\min(j_S,j_T)]$, then $\Phi_T' \term[i]$.
+  The result then follows easily by definition of $\lrol$ and $\lror$.
+\end{proof}
 

### 2018-06-18 fully-abstract-comp/technical-report.tex
made a claim more precise

@@
 
-    \lau{At this point we would like to use the IH, but we do not have $R_W'(a) \uplus W_M$ defined as $R_W'(a)$ is a split of $W_M$.}
+    \lau{At this point we would like to use the IH, but we do not have $R_W'(a) \uplus W_M$ defined because $R_W'(a)$ is a split of $W_M$ (in the case where the capability is $\linear$, in the other case we are fine.).}
 

### 2018-06-18 fully-abstract-comp/ftlr-proof-cases.tex
Compilation issue fixed

@@
   Our two languages are deterministic\dominique{Lemma needed for this, I guess..}, so we have that $\Phi_S \sterm{\ta,\stkb}$ iff $\Phi_S \sterm{\ta,\stkb}$ and $\Phi_T \term$ iff $\Phi_T' \term$.
-  It is also easy to show that if $\Phi_S \sterm[i + \min(j_S,j_T)]$, then $\Phi_S' \sterm[i]$ and if $\Phi_T \term[i+\min(j_S,j_T)]$, then $\Phi_T' \term[i]$.
+  It is also easy to show that if $\Phi_S \sterm[i + \min(j_S,j_T)]{\ta,\stkb}$, then $\Phi_S' \sterm[i]{\ta,\stkb}$ and if $\Phi_T \term[i+\min(j_S,j_T)]$, then $\Phi_T' \term[i]$.
   The result then follows easily by definition of $\lrol$ and $\lror$.

### 2018-06-18 fully-abstract-comp/technical-report.tex
removed ms_code from reasonability

@@
 \dominique{TODO: better name than ``reasonable'' for all of this.}
-Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
+Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
 We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if $n=0$
@@
 \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
+\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\ta)$
 \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
@@
 
-  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
+  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$:
+
     TODO
-  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
+  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint:
+
     This follows from $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$.

### 2018-06-18 fully-abstract-comp/technical-report.tex
Updated reasonability condition to have stk pointers in the reasonable words definition so it could be removed from the reasonable configuration definition.

@@
 Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
-We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if $n=0$
+We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
+free stack $\ms_\stk$ if $n=0$
 or the following implications hold.
@@
 \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
+\item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$
 \end{itemize}
@@
   \item $\Phi.\reg(\pcreg) = c$
-  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$
-  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
+  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$for all $r \not\in \{\pcreg\}$
   \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
@@
     TODO
+
+  \item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$:
+
+    TODO
+
   \end{itemize}
@@
   \begin{itemize}
-  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$
-  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
+  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$for all $r \not\in \{\pcreg\}$
   \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
@@
 
-  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$:
-
-    TODO
   \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint:

### 2018-06-18 fully-abstract-comp/ftlr-proof-cases.tex
Fix latex compilation.

--- 2018-06-18 fully-abstract-comp/ftlr-proof-cases.tex
@@
   Our two languages are deterministic\dominique{Lemma needed for this, I guess..}, so we have that $\Phi_S \sterm{\ta,\stkb}$ iff $\Phi_S \sterm{\ta,\stkb}$ and $\Phi_T \term$ iff $\Phi_T' \term$.
-  It is also easy to show that if $\Phi_S \sterm[i + \min(j_S,j_T)]$, then $\Phi_S' \sterm[i]$ and if $\Phi_T \term[i+\min(j_S,j_T)]$, then $\Phi_T' \term[i]$.
+  It is also easy to show that if $\Phi_S \sterm{i + \min(j_S,j_T)}$, then $\Phi_S' \sterm{i}$ and if $\Phi_T \term[i+\min(j_S,j_T)]$, then $\Phi_T' \term[i]$.
   The result then follows easily by definition of $\lrol$ and $\lror$.

### 2018-06-19 fully-abstract-comp/technical-report.tex
Corollary about pure worlds
Updated safety->reasonable lemma to require pure compatible worlds
typo

@@
 
+\begin{lemma}
+  \label{lem:comp-world-pure-comp-world}
+  If $W_1 \oplus W_2$ is defined, then $\purePart{W_1} \oplus \purePart{W_2}$ is defined
+\end{lemma}
+\begin{proof}
+  Follows from Lemma~\ref{lem:purePart-oplus}.
+\end{proof}
+
 \subsection{Memory satisfaction}
@@
   \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
-  \item $W_w \oplus W_M$ is defined
+  \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
   \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
@@
   \end{itemize}
-  then, with respect to $\code,\ta,\gsigrets,\gsigcloss$,
-  $w$ is reasonable up to $n$ steps in memory $\Phi.\mem$.
+  then, with respect to $\code,\ta,\gsigrets,\gsigcloss$, $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
 \end{lemma}
@@
   \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
-  \item $W_w \oplus W_M$ is defined
+  \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
   \item $\pwheap[W_w](r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$

### 2018-06-19 fully-abstract-comp/technical-report.tex
removed less than in the standard region and the code region.
Fix a leftover n' in the standard region's definition.
Remove comment that became outdated for the second time :).
Adjusted step in static region definition and typo in code region def.
Added the last case of lemma 20.

@@
       \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in A, n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \quad \forall \aaddr \in A \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
       \exists S : \dom(\ms) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in \dom(\ms) \ldotp \npair[n-1]{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \quad \forall \aaddr \in \dom(\ms) \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
-      \forall a \in \dom(\code), n' < n \ldotp\\
-      \quad\npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
+      \forall a \in \dom(\code)\ldotp\\
+      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
     \end{array}
@@
     In the second case, we have a contradiction with $\sigma \not\in (\gsigrets\cup \gsigcloss)$.
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_S(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
 
@@
 
-    From $\npair[n]{[b,e]} \in \readCond{\perm,\lin}$ we get $S\subseteq
-    \addressable{\lin,W}$ and $R : S \fun \powerset{\nats}$ such that
+    From $\npair[n]{[b,e]} \in \readCond{\perm,\lin}$ we get $S\subseteq \addressable{\lin,\pwheap}$ and $R : S \fun \powerset{\nats}$ such that
     \begin{itemize}
@@
 
-    By $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, we get $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that $\ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r))$ and $\ms_S = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r))$.
+    By $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, we get $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that $\ms_S = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r))$.
 
@@
 
-    We know $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$, so $\npair[n']{R_\ms(r)} \in \stdreg{R(r),\gc}{\pur}.H \; \xi^{-1}(R_W(r))$ for $n' < n$.
-    This means that for $n'' < n'$, we have $\npair[n'']{\pi_1(R_\ms(r))(a)} \in \lrv(R_W'(a))$ where $R_W' : [b,e] \fun \World$ such that $\biguplus_{a \in \dom(\pi_1(R_\ms(r)))} R_W'(a) = R_W(r)$.
-
-    \lau{At this point we would like to use the IH, but we do not have $R_W'(a) \uplus W_M$ defined because $R_W'(a)$ is a split of $W_M$ (in the case where the capability is $\linear$, in the other case we are fine.).}
-
-    TODO
+    We know $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$, so $\npair[n']{R_\ms(r)} \in \stdreg{R(r),\gc}{\pur}.H(\xi^{-1}(R_W(r)))$ for $n' < n$.
+    This gives us $\npair[n']{(\pi_1(R_\ms(r))(a),\_)} \in \lrv(R_W'(a))$ where $R_W' : [b,e] \fun \World$ such that $\biguplus_{a \in \dom(\pi_1(R_\ms(r)))} R_W'(a) = R_W(r)$.
+
+    At this point we apply the induction hypothesis which is possible as we have
+    the following:
+    \begin{itemize}
+    \item $\npair[n-1]{(\pi_1(R_\ms(r))(a),\_)} \in \lrv(R_W'(a))$
+
+      We get this by the above.
+    \item $\memSat[n-1]{\ms_S,\stk,\ms_\stk,\_}{W_M}$
+
+      This follows by assumption and Lemma~\ref{lem:downwards-closed}
+    \item $\purePart{R_W'(a)} \oplus \purePart{W_M}$
+
+      Which follows by definition of $\purePart{}$ and the fact that $R_W'(a)$ is $W_M$ with part of its ownership, but $\purePart{}$ strips away the ownership making the two compatible.
+    \item $\pwheap[W_w](r_\code) \nequal[n-1] \codereg{\sigrets,\sigcloss,\code,\gc}$
+
+      By property of $n$-equality.
+    \item $\dom(\code) \subseteq \ta$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
+
+      Follows by assumption.
+    \end{itemize}
+    which gives us that, with respect to $\code,\ta,\gsigrets,\gsigcloss$, $\pi_1(R_\ms(r))(a)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
+    Which is what we wanted as $\pi_1(R_\ms(r))(a) = \ms_S(a)$.
 
@@
 
-    TODO
-
+    This case is proven in the same way the previous case was.
+    The main difference is that the free part of the world is used instead of the heap part:
+
+    From $\npair{(w,\_)} \in \lrv(W_w)$ and $\perm \in \readAllowed{}$, we get $\npair{[b,e]} \in \stackReadCond{\perm}$.
+
+    From $\npair{[b,e]} \in \stackReadCond{\perm}$ we get $S\subseteq
+    \addressable{\linear,\pwfree}$ and $R : S \fun \powerset{\nats}$ such that
+    \begin{itemize}
+    \item $\biguplus_{r\in S} R(r) \supseteq [b,e]$
+    \item $\forall r\ldotp |R(r)|  = 1$
+    \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
+    \end{itemize}
+    given $a \in [b,e]$ we know by the above that there exists $r$ such that $a \in R(r)$.
+
+    By $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, we get $R_\ms : \dom(\activeReg{\pwfree}) \fun \MemSeg \times \MemSeg$ such that $\ms_\stk = \biguplus_{r \in \dom(\activeReg{\pwfree})} \pi_1(R_\ms(r))$.
+
+    Further, we get $R_W : \dom(\activeReg{\pwfree}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwfree})} R_W(r)$ and $\forall r \in \dom(\activeReg{\pwfree}) \ldotp \npair{R_\ms(r)} \in \pwfree(r).H \; \xi^{-1}(R_W(r))$.
+    For $W_M = W' \oplus W''$ for some $W''$.
+
+    In particular, we have $\npair{R_\ms(r)} \in \pwfree(r).H \; \xi^{-1}(R_W(r))$.
+
+    We know $\pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$, so $\npair[n']{R_\ms(r)} \in \stdreg{R(r),\gc}{\pur}.H(\xi^{-1}(R_W(r)))$ for $n' < n$.
+    This gives us $\npair[n']{(\pi_1(R_\ms(r))(a),\_)} \in \lrv(R_W'(a))$ where $R_W' : [b,e] \fun \World$ such that $\biguplus_{a \in \dom(\pi_1(R_\ms(r)))} R_W'(a) = R_W(r)$.
+
+    At this point we apply the induction hypothesis which is possible as we have
+    the following:
+    \begin{itemize}
+    \item $\npair[n-1]{(\pi_1(R_\ms(r))(a),\_)} \in \lrv(R_W'(a))$
+
+      We get this by the above.
+    \item $\memSat[n-1]{\ms_S,\stk,\ms_\stk,\_}{W_M}$
+
+      This follows by assumption and Lemma~\ref{lem:downwards-closed}
+    \item $\purePart{R_W'(a)} \oplus \purePart{W_M}$
+
+      Which follows by definition of $\purePart{}$ and the fact that $R_W'(a)$ is $W_M$ with part of its ownership, but $\purePart{}$ strips away the ownership making the two compatible.
+    \item $\pwheap[W_w](r_\code) \nequal[n-1] \codereg{\sigrets,\sigcloss,\code,\gc}$
+
+      By property of $n$-equality.
+    \item $\dom(\code) \subseteq \ta$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
+
+      Follows by assumption.
+    \end{itemize}
+    which gives us that, with respect to $\code,\ta,\gsigrets,\gsigcloss$, $\pi_1(R_\ms(r))(a)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
+    Which is what we wanted as $\pi_1(R_\ms(r))(a) = \ms_\stk(a)$.
   \end{itemize}

### 2018-06-19 fully-abstract-comp/technical-report.tex
Prove lemma that is a simple consequence of another.

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
   $\oplus$ is associative and commutative.
+  Also, left-hand sides in the commutativity and associativity laws are defined whenever the right-hand sides are defined and vice versa.
 \end{lemma}
@@
   For all $W_1,W_2$, we have that $\purePart{W_1\oplus W_2} = \purePart{W_1} \oplus \purePart{W_2}$.
+  Also, the latter is defined when the former is defined (but not necessarily vice versa).
 \end{lemma}
@@
       \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in A, n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \quad \forall \aaddr \in A \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
 \lau{01-11-2017: RESOLVED look into what kind of world separation should happen here.}
-%\dominique{7-3-2018: is it necessary to have $n-1$ above?
-there is already a quantification over $n'<n$ in the memory satisfaction relation.} \lau{04-04-2018: The n'<n was removed from memory satisfaction.}
 
@@
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
-      \forall a \in \dom(\code), n' < n \ldotp\\
+      \forall a \in \dom(\code)\ldotp\\
       \quad\npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
@@
 
-\begin{lemma}
+\begin{lemma}[Untrusted source values are reasonable]
+  \label{lem:untrusted-source-values-are-reasonable}
   If
@@
   \begin{itemize}
+  \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
   \item $\npair{(\reg_S,\_)} \in \lrr(W_w)$
@@
   \begin{itemize}
-  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$for all $r \not\in \{\pcreg\}$
-  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
+  \item $\reg_S(r)$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$ for all $r \not\in \{\pcreg\}$
+  \item $\ms_S$, $\ms_\stk$ and $\stk$ are all disjoint
   \end{itemize}
@@
   \begin{itemize}
-  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$:
-
-    We have that $W = \bigoplus_{r \in \RegName \setminus \{\pcreg\}} S(r)$ and $\npair{(\reg_S(r),\reg_T(r))} \in \lrv(S(r))$ for all $r \not\in \{\pcreg\}$.
-
-    TODO
-
-  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint:
+  \item $\reg_S(r)$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$ for all $r \not\in \{\pcreg\}$:
+
+    We have that $W_w = \bigoplus_{r \in \RegName \setminus \{\pcreg\}} S(r)$ and $\npair{(\reg_S(r),\reg_T(r))} \in \lrv(S(r))$ for all $r \not\in \{\pcreg\}$.
+
+    For all $r \neq \pcreg$, we have that $\purePart{S(r)} \oplus \purePart{W_M}$ is defined by Lemmas~\ref{lem:purePart-oplus} and~\ref{lem:oplus-assoc-comm}.
+
+    The result then follows directly from Lemma~\ref{lem:untrusted-source-values-are-reasonable}.
+
+  \item $\ms_S$, $\ms_\stk$ and $\stk$ are all disjoint:
 

### 2018-06-19 fully-abstract-comp/technical-report.tex

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
       \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in A \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \quad \forall \aaddr \in A \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}

### 2018-06-19 fully-abstract-comp/technical-report.tex
Prove that reasonable things don't need to be trusted, fix lrp along the way.

@@
     \begin{array}{l}
-      \exists r \in \addressable{l,W} \ldotp\\
-      \quad \forall \npair{\stpair{\ms}{\ms}} \in S(i)(r).H \; \xi^{-1}(W)  \ldotp \\
-      \qquad \dom(\src{\ms_S}) \supseteq [\baddr,\eaddr] \wedge \dom(\ms_T) \supseteq [\baddr,\eaddr]
+      \exists S \subseteq \addressable{l,W} \ldotp\\
+      \exists P : S \rightarrow \powerset{\Addr} \ldotp\\
+      \quad \uplus_{r \in S} P(r) \supseteq [\baddr,\eaddr] \\
+      \quad \forall r \in S, W'\ldotp \npair{\stpair{\ms}{\ms}} \in \pwheap(r).H \; W'  \ldotp \\
+      \qquad \dom(\src{\ms_S}) = P(r) \wedge \dom(\ms_T) = P(r)
     \end{array}
@@
 
-\begin{lemma}
+\begin{lemma}[Untrusted register files are reasonable]
   If
@@
 
+\begin{lemma}[Reasonable things don't need to be trusted]
+  \label{lem:trusted-and-reasonable-is-untrusted}
+  If
+  \begin{itemize}
+  % \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
+  \item $\npair{(w,\_)} \in \lrvg{\trusted}(W_w)$
+  % \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
+  % \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
+  % \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
+  % \item $\dom(\code) \subseteq \ta$
+  % \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
+  \item $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$, with respect to $\code,\ta,\gsigrets,\gsigcloss$.
+  \item $\gsigrets \neq \emptyset$
+  \item $n > 0$
+  \item Theorem~\ref{thm:ftlr} holds up to $n$ steps.
+    \dominique{mention this assumption explicitly to avoid cyclic reasoning in the FTLR proof.}
+  \end{itemize}
+  Then
+  \begin{itemize}
+  \item $\npair{(w,\_)} \in \lrv(W_w)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Assume that $\npair{(w,w_T)} \in \lrvg{\trusted}(W_w)$
+
+  By definition of $\lrvg{\trusted}$, it is clear that also $\npair{(w,w_T)} \in \lrv(W_w)$, except in the following two cases:
+  \begin{itemize}
+  \item
+    \begin{itemize}
+    \item $w = \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}}$
+    \item $w_T = \seal{\sigma,\sigma_\baddr,\sigma_\eaddr}$ 
+    \item $r \in \dom(\pwheap)$
+    \item $\pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
+    \item $\dom(\code) \subseteq \ta$ 
+    \item $[\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)$ 
+    \item $\sigrets \subseteq \gsigrets$
+    \item $\sigcloss \subseteq \gsigcloss$
+    \end{itemize}
+
+    By definition of reasonability, we get immediately that $[\sigma_b,\sigma_e]\mathrel{\#} (\gsigrets\cup \gsigcloss)$, contradicting the last three facts above.
+  \item 
+    \begin{itemize}
+    \item $w = \src{((\perm,\normal),\baddr,\eaddr,\aaddr)}$, 
+    \item $w_T =
+      ((\perm,\normal),\baddr,\eaddr,\aaddr)$
+    \item $\perm \sqsubseteq \rx$
+    \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
+    \item $\npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W}$
+    \end{itemize}
+
+    If $\perm = \noperm$, then $\npair{(w,w_T)} \in \lrv(W_w)$ follows if we can prove that $\npair{(\normal,b,e)} \in \lrp(W)$.
+    This follows easily by definition of $\lrp$, $\xReadCond{}$ and $\codereg{}$.
+
+    If $\perm \neq \noperm$, then by definition of reasonability, we get immediately that $[b,e] \mathrel{\#} \ta$.
+
+    By definition of $\lrv$, it suffices to show that $\npair{[b,e]} \in \readCond{\normal,W}$ and $\npair{[b,e]} \in \execCond{\normal,W}$.
+    The former follows directly from Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
+    The latter follows by Theorem~\ref{thm:ftlr}, Lemma~\ref{lem:conds-shrinkable} and definition of $\execCond{}$.
+  \end{itemize}
+\end{proof}
+
 \subsection{Fundamental Theorem of Logical Relations}
@@
   \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
+    \dominique{This requirement is actually not needed because of Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}}
   \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
@@
   Follows by defintion of $\readCond{}$ from Lemma~\ref{lem:codereg-untrusted-stdreg}.
+\end{proof}
+
+\begin{lemma}[Code memory for untrusted components safely readable]
+  \label{lem:xReadCond-outside-ta-implies-readCond}
+  If
+  \begin{itemize}
+  \item $A \mathrel{\#} \ta$
+  \item $\npair{A} \in \xReadCond{W}$
+  \end{itemize}
+  we have that $\npair{A} \in \readCond{\normal,W}$.
+\end{lemma}
+\begin{proof}
+  Follows by defintion of $\readCond{}$ and $\xReadCond{}$ from Lemma~\ref{lem:codereg-untrusted-stdreg}.
 \end{proof}

### 2018-06-19 fully-abstract-comp/technical-report.tex
Definition environments for reasonability definitions.

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
 or the following implications hold.
-\begin{itemize}
-\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\ta)$
-\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
-\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
-\item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$
-\end{itemize}
-
-We say that an execution configuration $\Phi$ is reasonable up to $n$ steps if:
+\begin{definition}[Reasonable word]
+  \label{def:reasonable-word}
+  \begin{itemize}
+  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then
+    $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets
+    \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\ta)$
+  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
+    \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$
+    and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all
+    $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
+  \item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$
+    and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for
+    all $\aaddr \in [\baddr,\eaddr]$
+  \end{itemize}
+\end{definition}
+
+\begin{definition}[Reasonable configuration]
+  \label{def:reasonable-conf}
+  We say that an execution configuration $\Phi$ is reasonable up to $n$ steps
+  if:
   \begin{itemize}
@@
     \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
+    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$
+      for some $\src{r_1}$ and $\src{r_2}$
     \item $0 < i \le n$
@@
     \begin{itemize}
-    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
+    \item $\src{\Phi'}(\src{r_\stk}) =
+      \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
+    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to
+      $n-i$ steps
     \end{itemize}
@@
     \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
+    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and
+      $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
     \item $0 < i \le n$
@@
     \begin{itemize}
-    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$ 
+    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$
+      and $\sigma \in \gsigrets$
     \item $\sigma \in \gsigcloss$.
@@
     \end{itemize}
-    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
-  \end{itemize}
+    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for
+    all $r$ (including $\pcreg$)
+  \end{itemize}  
+\end{definition}
 
@@
 
-We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps if for any $\Phi$ such that
+  \begin{definition}[Reasonable pc]
+    \label{def:reasonable-pc}
+    We say that an executable capability $c =
+    ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps
+    if for any $\Phi$ such that
     \begin{itemize}
     \item $\Phi.\reg(\pcreg) = c$
-  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$for all $r \not\in \{\pcreg\}$
+    \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and
+      free stack $\Phi.\ms_\stk$for all $r \not\in \{\pcreg\}$
     \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
@@
     We have that $\Phi$ is reasonable.
-
-We say that a component $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$ is reasonable if the following hold:
-For all $\sealed{\sigma,\vsc} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$, we have that $\vsc$ behaves reasonably up to any number of steps $n$.
-\lau{This need to change with the proposed simplification of the reasonability update.}
+  \end{definition}
+
+  \begin{definition}[Reasonable component]
+    \label{def:reasonable-component}
+    We say that a component
+    $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$
+    is reasonable if the following hold: For all $\sealed{\sigma,\vsc} \in
+    \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$, we have that $\vsc$
+    behaves reasonably up to any number of steps $n$.
+    \lau{Is this reasonable as a word? In that case, it should  mention the
+      memory and free stack.}
+  \end{definition}
 

### 2018-06-19 fully-abstract-comp/ftlr-proof-cases.tex
Technical fixes + remove unnecessary requirement in FTLR.

--- 2018-06-18 fully-abstract-comp/ftlr-proof-cases.tex
@@
 \begin{proof}[Proof of Theorem~\ref{thm:ftlr}]
+  Note first that if $\trust = \untrusted$, then $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ follows by Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
+
   By general induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.

### 2018-06-19 fully-abstract-comp/technical-report.tex
Technical fixes + remove unnecessary requirement in FTLR.

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
   \begin{itemize}
-  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
-    \dominique{This requirement is actually not needed because of Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}}
   \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
@@
   \item $\dom(\mscode) \subseteq \ta \vee (\dom(\mscode)\mathrel{\#} \ta \wedge \sigrets = \emptyset)$, but we know that $\emptyset \neq\dom(\mscode) \mathrel{\#} \ta$, so we get that $\sigrets = \emptyset$.
-  \item For all $a \in \dom(\code)$, $n'<n$, $\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$
+  \item For all $a \in \dom(\code)$, $\npair{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$
   \end{itemize}
@@
   \item $\dom(\mscode) = \dom(\mscode) = \dom(\mscode)$
-  \item there exists a $S : \dom(\mscode) \fun \World$ with $\xi(\hat{W}) = \oplus_{\aaddr\in\dom(\mscode)} S(\aaddr)$ and for all $\aaddr \in \dom(\mscode)$, $n'<n$, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$
+  \item there exists a $S : \dom(\mscode) \fun \World$ with $\xi(\hat{W}) = \oplus_{\aaddr\in\dom(\mscode)} S(\aaddr)$ and for all $\aaddr \in \dom(\mscode)$, we have that $\npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$
   \end{itemize}
@@
   We take $S$ to map every address to $\purePart{\hat{W}}$, except for a single one that we map to $\hat{W}$.
-  The result then follows by Lemmas~\ref{lem:purePart-duplicable}, \ref{lem:oplus-future} and \ref{lem:monotonicity} and the above fact that for all $a \in \dom(\code)$, $n'<n$, $\npair[n']{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$.
+  The result then follows by Lemmas~\ref{lem:purePart-duplicable}, \ref{lem:oplus-future} and \ref{lem:monotonicity} and the above fact that for all $a \in \dom(\code)$, $\npair{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$.
 \end{proof}

### 2018-06-19 fully-abstract-comp/technical-report.tex
Clarify comment.
Made ce reasonable definition downwards closed.

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
   We say that an execution configuration $\Phi$ is reasonable up to $n$ steps
-  if:
+  iff for all $n' \le n$:
   \begin{itemize}
@@
       for some $\src{r_1}$ and $\src{r_2}$
-    \item $0 < i \le n$
+    \item $0 < i \le n'$
     \end{itemize}
@@
     \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to
-      $n-i$ steps
+      $n'-i$ steps
     \end{itemize}
@@
     \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and
-      $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
-    \item $0 < i \le n$
+    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
     \end{itemize}
@@
     \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
-    \item $0 < i \le n$
-    \end{itemize}
-    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
+    \item $0 < i \le n'$
+    \end{itemize}
+    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n'-i$ steps
   \item \emph{Don't leak private stuff...} If
@@
     \item $\src{\Phi''}$ is not in $\ta$
-    \item $0 < i \le n$
-    \end{itemize}
-    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for
+    \item $0 < i \le n'$
+    \end{itemize}
+    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n'-i$ steps for
     all $r$ (including $\pcreg$)
@@
     Follows easily by definition.
+  \end{proof}
+
+  \begin{lemma}
+    \label{lem:ec-reasonable-downwards-closed}
+    For all $n' \le n$ if
+    \begin{itemize}
+    \item $\Phi$ is reasonable up to $n$ steps
+    \end{itemize}
+    Then
+    \begin{itemize}
+    \item $\Phi$ is reasonable up to $n'$ steps
+    \end{itemize}
+  \end{lemma}
+  \begin{proof}
+    Follows by definition.
   \end{proof}

### 2018-06-19 fully-abstract-comp/technical-report.tex

@@
   \item Theorem~\ref{thm:ftlr} holds up to $n$ steps.
-    \dominique{mention this assumption explicitly to avoid cyclic reasoning in the FTLR proof.}
+    \dominique{Note: I mention this assumption explicitly to avoid cyclic reasoning in the FTLR proof.}
   \end{itemize}

### 2018-06-19 fully-abstract-comp/ftlr-proof-cases.tex
typo
Looked through case 3 of FTLR proof and one reformulation.

@@
     \begin{itemize}
-    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
-    \end{itemize}:
-    Follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep{1} \Phi_S'$ with Lemma~\ref{lem:ec-reasonable-antired}.
-    \dominique{not actually true: reasonability definitions need to be made downward closed using quantification over smaller $n'$?}
-  \item $\trust = \trusted \vee \npair[n'-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$:
-    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}.
-  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$:
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n'-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $\npair[n'-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \end{itemize}: Follows from the corresponding assumption of this lemma and
+    the fact that $\Phi_S \nstep[1]{\ta,\stkb} \Phi_S'$ and
+    Lemma~\ref{lem:ec-reasonable-antired},
+    \ref{lem:ec-reasonable-downwards-closed}, and \ref{lem:downwards-closed}.
+  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$:
     Follows from the definition of $\updPcAddr{}$ and the assumptions of this case.

### 2018-06-19 fully-abstract-comp/technical-report.tex
Looked through case 3 of FTLR proof and one reformulation.
Minor unfinished work on a component-related proof.

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
 
+    First, we have that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$ for all $n' < n$.
+    This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
+
     By definition of $\lrv$ and by choosing island $r = r_{\mathrm{code}}$, it suffices to prove that
@@
      \end{itemize}
-     Since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$ in both cases.
-     This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$), we know that $\npair[n']{(\vsc,\vsc)} \in \lrv(W)$.
+     Both follow since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$, $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, and the above fact that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$.
 
@@
      \end{equation*}
+
+     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, we know that
+     one of the following holds:
+     \begin{itemize}
+     \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrv \; \xi(\xi^{-1}(W_o))$
+     \item $(\dom(\code) \subseteq \ta$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrvtrusted \; \xi(\xi^{-1}(W_o))$
+     \end{itemize}
+
      \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}

### 2018-06-20 fully-abstract-comp/technical-report.tex

--- 2018-06-19 fully-abstract-comp/technical-report.tex
@@
   \begin{proof}
-    Follows by definition.
+    Follows from the definition.
   \end{proof}

### 2018-06-20 fully-abstract-comp/technical-report.tex
layout improvement.
Latex fix, remove duplicate requirement.
Some changes needed to prove a remaining lemma from the component related lemmas.
The required changes include the requirement that non-executable capabilities
signed with closure seals must be in the *untrusted* value relation.
Made some corresponding changes throughout the document.

@@
   \inference{
-    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} & 
-    \sigma \in \sigcloss
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
-  }
-  \and
-  \inference{
     \permbnf \sqsubseteq \rw&
@@
   \inference{
+    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} & 
+    \sigma \in \sigcloss
+  }{
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-export}} \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
+  }
+  \and
+  \inference{
+    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\emptyset,\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+  }{
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w
+  }
+\end{mathpar}
+
+\begin{mathpar}
+  \inference{
     \dom(\mscode) \mathrel{\#} \dom(\msdata)&
@@
     \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
-    \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}\\
+    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} c_{\mathrm{export}}}\\
     \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)
@@
       \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
+      {} [\baddr,\eaddr] \subseteq \ta \wedge\\
       \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
@@
 
-\begin{equation*}
-  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
+\begin{multline*}
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
+  \left\{\npair{(\code,\code)} \middle|
     \begin{array}{l}
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \trust = \trusted) \vee \\
+      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \trust = \trusted) \vee \\
       \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
@@
   \right\}
-\end{equation*}
+\end{multline*}
 
@@
       \begin{array}{l}
-        \sigma \in \sigcloss \tand\\
         (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
+        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
+         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
+         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
       \end{array}
@@
     \end{itemize}
-  \item \emph{Use return seals only for calls} If
+  \item \emph{Use return seals only for calls, use closure seals appropriately} If
     \begin{itemize}
     \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item i > 0
     \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
@@
       and $\sigma \in \gsigrets$
-    \item $\sigma \in \gsigcloss$.
+    \item $\sigma \in \gsigcloss$ and one of the following holds:
+      \begin{itemize}
+      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n$ steps.
+      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
+      \end{itemize}
     \end{itemize}
@@
     \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and
-      free stack $\Phi.\ms_\stk$for all $r \not\in \{\pcreg\}$
+      free stack $\Phi.\ms_\stk$ for all $r \not\in \{\pcreg\}$
     \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
@@
 Note: we don't require the readcondition in the trusted case because trusted code pointers point to trusted code blocks which may require seals for trusted seals which are not in the untrusted value relation.
+
+\begin{lemma}[Untrusted environments produce safe closures]
+  \label{lem:lre-implies-lrexj}
+  If
+  \begin{itemize}
+  \item $\npair{(w_1,w_1')} \in \lre(W_1)$ or ($\nonExec{w_1}$ and $\nonExec{w_1'}$)
+  \item $w_1 \neq \retptrc(\_)$
+  \item $\npair{(w_2,w_2')} \in \lrv(W_2)$ or ($\exec{w_2}$ and $\exec{w_2'}$)
+  \end{itemize}
+  Then
+  \begin{itemize}
+  \item $\npair{((w_1,w_2),(w_1',w_2'))} \in \lrexj(W_1 \oplus W_2)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Take $n' \leq n$, $\src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}$, $W_\lrrs , W_\lrm$ and assume
+  \begin{itemize}
+  \item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\untrusted}(\{\rdata\}) (W_\lrrs )$
+  \item $\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm}$
+  \item $\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}$
+  \item $\Phi_T = (\ms_T,\reg_T)$
+  \item $W_1 \oplus W_2 \oplus W_\lrrs \oplus W_\lrm$ is defined 
+  \end{itemize}
+  Then we need to prove that for $\Phi_S'$,$\Phi_T'$
+  \begin{itemize}
+  \item $\Phi_S' = \xjumpResult{w_1}{w_2}{\Phi_S}$
+  \item $\Phi_T' = \xjumpResult{w_1'}{w_2'}{\Phi_T}$
+  \item $\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro$
+  \end{itemize}
+
+  Using the fact that $w_1 \neq \retptrc(\_)$, we know by definition of $\xjumpResult{w_1}{w_2}{\Phi_S}$ that we must be in one of the following two cases:
+  \begin{itemize}
+  \item $w_1 \neq \retptrc(\_)$, $w_2 \neq \retptrd(\_)$, $\nonExec{w_2}$ and
+    $\Phi_S' = \Phi_S \updReg{\pcreg}{w_1}\updReg{\rdata}{w_2}$.
+
+    From $\nonExec{w_2}$ and $\npair{(w_2,w_2')} \in \lrv(W_2)$ or $\exec{w_2}$, it follows that also $\nonExec{w_2'}$ and $\Phi_T' = \Phi_T \updReg{\pcreg}{w_1'}\updReg{\rdata}{w_2'}$
+
+    We can now combine $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs )$ and $\npair{(w_2,w_2')} \in \lrv(W_2)$ into $\npair[n']{(\reg_S\update{\rdata}{w_2},\reg_T\update{\rdata}{w_2'})} \in \lrr (W_\lrrs \oplus W_2)$, using Lemmas~\ref{lem:monotonicity},~\ref{lem:oplus-future},~\ref{lem:oplus-assoc-comm}, and~\ref{lem:downwards-closed}.
+
+    With our other assumptions above, $\npair{(w_1,w_1')} \in \lre(W_1)$ then gives us that $\npair[n']{(\Phi_S',\Phi_T')}\in \lro$, as required.
+    On the other hand, if $\nonExec{w_1}$ and $\nonExec{w_1'}$, then $\Phi_S' \step[\gc] \failed$ and $\Phi_T' \step \failed$ and the result follows by definition of $\lrol$ and $\lror$.
+
+  \item Otherwise: $\Phi_S' = \failed$.
+
+    From $w_1 \neq \retptrc(\_)$ and ($\npair{(w_2,w_2')} \in \lrv(W_2)$ or ($\exec{w_2}$ and $\exec{w_2'}$)), the only way we can get to this case is that $\exec{w_2}$ and $\exec{w_2'}$.
+
+    It follows that $\Phi_T' = \failed$.
+
+    The result follows by definition of $\lrol$ and $\lror$.
+  \end{itemize}
+\end{proof}
+
+\begin{lemma}[Safe values are safe to execute]
+  \label{lem:safe-vals-safe-exec}
+  If
+  \begin{itemize}
+  \item $\npair{(w_1,w_1')} \in \lrv(W_1)$
+  \item $\npair[n']{(w_2,w_2')} \in \lrv(W_2)$ or ($\exec{w_2}$ and $\exec{w_2'}$)
+  \item $n' < n$
+  \end{itemize}
+  Then
+  \begin{itemize}
+  \item $\npair[n']{((w_1,w_2),(w_1',w_2'))} \in \lrexj(W_1 \oplus W_2)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  By Lemma~\ref{lem:lre-implies-lrexj}, it suffices to prove the following: 
+  \begin{itemize}
+  \item $\npair[n']{(w_1,w_1')} \in \lre(W_1)$ or ($\nonExec{w_1}$ and $\nonExec{w_1'}$)
+  \item $w_1 \neq \retptrc(\_)$.
+  \end{itemize}
+
+  The latter follows immediately from $\npair{(w_1,w_1')} \in \lrv(W_1)$ by definition.
+  It also follows that either ($\nonExec{w_1}$ and $\nonExec{w_1'}$) or $w_1 = w_1' = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ and $\npair{[\baddr,\eaddr]} \in \execCond{W_1}$.
+  In the latter case, it follows by definition of $\execCond{}$ that $\npair[n']{(w_1,w_1')} \in \lre(W_1)$ as required.
+\end{proof}
 
@@
   \item $w = c$, $c \in \overline{c_{\mathrm{import}}}$: contradicts assumption.
-  \item $w = \sealed{\sigma,\vsc}$, $\vsc = ((\rx,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq \dom(\mscode)$, $\sigma \in \sigcloss$:
-
-    By definition of $\lrv$ and by choosing island $r = r_{\mathrm{code}}$, it suffices to prove that
-    \begin{itemize}
-    \item $\npair[n']{(\vsc,\vsc)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
-
-     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that one of the following holds:
-     \begin{itemize}
-     \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrv \; \xi(\xi^{-1}(W))$
-     \item $(\dom(\code) \subseteq \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\xi^{-1}(W))$
-     \end{itemize}
-     Take $\trust = \untrusted$ iff $\dom(\mscode \mathrel{\#} \ta)$ and $\trust = \trusted$ iff $\dom(\mscode) \subseteq \ta$.
-     Since $\xi(\xi^{-1}(W)) = W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrvg{\trust}(W)$.
-
-     This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs}, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
-
-   \item $(\isLinear{\vsc} \text{ iff } \isLinear{\vsc})$: trivially fine
-
-   \item 
-     For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$, we have that
-     \begin{equation*}
-       \npair[n']{(\vsc,\src{\vsc_S'},\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
-     \end{equation*}
-
-     Since $\sigma\in\sigcloss$, it follows from $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$ that
-     \begin{equation*}
-       \npair[n']{(\vsc_S',\vsc'_T)} \in \lrvg{\trust}(W_o)
-     \end{equation*}
-     with $\trust = \untrusted$ iff $\dom(\code \mathrel{\#} \ta)$ and $\trust = \trusted$ iff $\dom(\code) \subseteq \ta$.
-
-     As above, it follows from Lemma~\ref{lem:ftlr-component-code-ptrs}, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$ that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(\purePart{W})$.
-
-     \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
-     TODO
-   \end{itemize}
-
   \item $w = ((\perm,\lin),\baddr,\eaddr,\aaddr)$,
@@
 
-    First, we have that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$ for all $n' < n$.
-    This follows by induction and Lemma~\ref{lem:downwards-closed} (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
+    First, we have that $\npair{(\vsc,\vsc)} \in \lrv (W)$.
+    This follows by induction (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
 
@@
      \end{itemize}
-     Both follow since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$, $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, and the above fact that $\npair[n']{(\vsc,\vsc)} \in \lrv (W)$.
+     Both follow since $\xi(\xi^{-1}(W)) = W$ and $\lrv \; W \subseteq \lrvtrusted \; W$, $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, and the above fact that $\npair{(\vsc,\vsc)} \in \lrv (W)$ and Lemma~\ref{lem:downwards-closed}.
 
@@
      \begin{equation*}
-       \npair[n']{(\vsc,\vsc_S',\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o)) \wedge \\
+       \npair[n']{(\vsc,\vsc_S',\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
      \end{equation*}
@@
      \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrv \; \xi(\xi^{-1}(W_o))$
-     \item $(\dom(\code) \subseteq \ta$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrvtrusted \; \xi(\xi^{-1}(W_o))$
-     \end{itemize}
-
-     \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
+     \item $(\dom(\code) \subseteq \ta$ and $\sigcloss\subseteq \gsigcloss$ and $\sigrets \subseteq \gsigrets$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrvg{\trust} \; \xi(\xi^{-1}(W_o))$ with $\trust = \trusted$ iff $\exec{\vsc_S'}$.
+     \end{itemize}
+
+     Lemma~\ref{lem:safe-vals-safe-exec} now allows us to conclude (using Lemma~\ref{lem:monotonicity}) that:
+     \begin{equation*}
+       \npair[n']{(\vsc,\vsc_S',\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
+     \end{equation*}
+
    \item 
@@
        \end{equation*}
-       \dominique{TODO:I expect this should be able to reuse a result that we also need for the regular FTLR?}
-    \end{itemize}
+
+     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, we know that
+     one of the following holds:
+     \begin{itemize}
+     \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrv \; \xi(\xi^{-1}(W_o))$
+     \item $(\dom(\code) \subseteq \ta$ and $\sigcloss\subseteq \gsigcloss$ and $\sigrets \subseteq \gsigrets$ and $\npair[n']{(\vsc_S',\vsc_T')} \in \lrvg{\trust} \; \xi(\xi^{-1}(W_o))$ with $\trust = \trusted$ iff $\exec{\vsc_S'}$.
+     \end{itemize}
+
+     First, Lemma~\ref{lem:non-linear-pure} allows us to conclude that also $\npair{(\vsc,\vsc)} \in \lrv (\purePart{W})$.
+
+     Next, Lemma~\ref{lem:safe-vals-safe-exec} now allows us to conclude (using Lemma~\ref{lem:monotonicity}) that:
+     \begin{equation*}
+       \npair[n']{(\vsc,\vsc_S',\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
+     \end{equation*}
+    \end{itemize}
+  \end{itemize}
+\end{proof}
+
+\begin{lemma}[FTLR for component exports]
+  \label{lem:ftlr-comp-export}
+  If
+  \begin{equation*}
+    \dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w
+  \end{equation*}
+  and
+  \begin{itemize}
+  \item $\npair{A_{\mathrm{non-linear}}} \in \readCond{\normal,W}$
+  \item $\npair{A_{\mathrm{non-linear}}} \in \writeCond{\normal,W}$
+  \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
+  \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$
+  \item If $w = \sealed{\sigma,\vsc}$ with $\sigma \in \sigcloss$ and $\exec{\vsc}$, then
+    $\vsc$ behaves reasonably up to $n$ steps.
+  \end{itemize}
+
+  then $\npair{(w,w)} \in \lrv(W)$
+\end{lemma}
+\begin{proof}
+  By induction on the judgement
+  \begin{equation*}
+    \dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w
+  \end{equation*}
+  We have the following cases:
+  \begin{itemize}
+  \item $w = \sealed{\sigma,\vsc}$, $\vsc = ((\rx,\normal),\baddr,\eaddr,\aaddr)$, $[\baddr,\eaddr]\subseteq \dom(\mscode)$, $\sigma \in \sigcloss$:
+
+    By definition of $\lrv$ and by choosing island $r = r_{\mathrm{code}}$, it suffices to prove that
+    \begin{itemize}
+    \item $\npair[n']{(\vsc,\vsc)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
+
+      By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that one of the following holds:
+      \begin{itemize}
+      \item $(\dom(\code) \mathrel{\#} \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrv \; \xi(\xi^{-1}(W))$
+      \item $(\dom(\code) \subseteq \ta$ and $\npair[n']{(\vsc,\vsc)} \in \lrvtrusted \; \xi(\xi^{-1}(W))$ and $\exec{\vsc}$
+      \end{itemize}
+      Take $\trust = \untrusted$ iff $\dom(\mscode \mathrel{\#} \ta)$ and $\trust = \trusted$ iff $\dom(\mscode) \subseteq \ta$.
+      Since $\xi(\xi^{-1}(W)) = W$ and because we know by assumption that $\dom(\mscode \mathrel{\#} \ta)$ or $\dom(\mscode) \subseteq \ta$, it suffices to prove that $\npair[n']{(\vsc,\vsc)} \in \lrvg{\trust}(W)$.
+
+      This last fact follows from Lemma~\ref{lem:ftlr-component-code-ptrs}, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+
+    \item $(\isLinear{\vsc} \text{ iff } \isLinear{\vsc})$: trivially fine
+
+    \item
+      For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$, we have that
+      \begin{equation*}
+        \npair[n']{(\vsc,\src{\vsc_S'},\vsc,\vsc_T')} \in \lrexj(W'\oplus W_o))
+      \end{equation*}
+
+      % As above, it follows from Lemma~\ref{lem:ftlr-component-code-ptrs}, using Lemma~\ref{lem:downwards-closed} and the definition of $\xReadCond{}$ with the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$ that $\npair{(\vsc,\vsc)} \in \lrvg{\trust}(\purePart{W})$.
+
+      From one of our assumptions, we know that $\vsc$ behaves reasonably up to $n$ steps.
+
+      Theorem~\ref{thm:ftlr} now tells us that $\npair{(\vsc,\vsc)} \in \lre(\purePart{W})$, using the definition of $\xReadCond{}$ and the assumption that $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$.
+
+      Since $\sigma\in\sigcloss$, it follows from $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$ that
+      \begin{equation*}
+        \npair[n']{(\vsc_S',\vsc'_T)} \in \lrvg{\trust}(W_o)
+      \end{equation*}
+      with $\trust = \untrusted$ iff ($\dom(\code \mathrel{\#} \ta)$ or $\nonExec{\vsc_S'}$) and $\trust = \trusted$ iff ($\dom(\code) \subseteq \ta$ and $\exec{\vsc_S'}$).
+
+      The result now follows from Lemma~\ref{lem:lre-implies-lrexj}.
+    \end{itemize}
+
+  \item TODO: the other case...
   \end{itemize}
@@
 
-    By Lemma~\ref{lem:ftlr-comp-value}, it then suffices to prove that
+    By Lemma~\ref{lem:ftlr-comp-export}, it then suffices to prove that
     \begin{itemize}
@@
     \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$: by assumption.
-    \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(\purePart{}(\purePart{W'}))$ for all $n' < n$: by the fact that $\purePart{}(\purePart{W'}) = \purePart{W'}$ (Lemma~\ref{lem:purePart-idempotent}) and the above assumption that
-      \begin{equation*}
-        \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$} 
-      \end{equation*}
-    \item $w \not\in \overline{c_{\mathrm{import}}}$: by assumption.
     \end{itemize}

### 2018-06-21 fully-abstract-comp/ftlr-proof-cases.tex
Fix some open todos.

@@
   \item $\memSat{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$
+  \item $W_\pcreg \oplus W_R \oplus W_M$ is defined.
   \item Theorem~\ref{thm:ftlr} holds for all $n' < n$.
@@
   TODO: update...
-  By general induction on $n$, i.e. we can assume that the lemma already holds for $n' < n$.
+  By complete induction on $n$, i.e. we can assume that the lemma already holds for $n' < n$.
 
@@
 \begin{proof}[Proof of Theorem~\ref{thm:ftlr}]
-  Note first that if $\trust = \untrusted$, then $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ follows by Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
-
-  By general induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
+  By complete induction over $n$\footnote{if $n=0$, then we have a contradiction with $\Phi_S\term[i]$ and $\Phi_T\term[i]$ when we get to $\lro$.}.
   Assume
   \begin{itemize}
-  \item $\trust = \trusted \vee \npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$
   \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
@@
     \item $[\baddr,\eaddr] \subseteq \ta$
-    \item $\trust = \trusted$
     \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
@@
     \item $[\baddr,\eaddr] \mathrel{\#} \ta$
-    \item $\trust = \untrusted$
     \end{itemize}
@@
 \[
-\npair{(c,c)} \in \lreg{\trust}(W)
+\npair{(c,c)} \in \lre(W)
 \]
@@
 \begin{enumerate}
-\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$ \label{item:reg-ass}
+\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$ \label{item:reg-ass}
 \item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
-\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
+\item $W \oplus W_R \oplus W_M$ is defined\label{item:worlds-def}
 \end{enumerate}
@@
   TODO
-\item $\trust = \trusted \vee \npair[n']{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$:
-  TODO
+
+\item $\trust = \trusted \vee \npair[n']{[\baddr,\eaddr]} \in \readCond{\lin,W}$:
+  If $\trust =\ untrusted$ then we know that $[b,e] \mathrel{\#} \ta$, so that $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ follows by Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
+
 \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr)$:
-  TODO
-\item $\npair[n']{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
-  TODO
+  We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c = ((\rx,\lin),\baddr,\eaddr,\aaddr)$.
+
+\item $\npair[n']{[\baddr,\eaddr]} \in \xReadCond{W}$:
+  Follows from $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ using Lemma~\ref{lem:downwards-closed}.
+
 \item $\npair[n']{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
-  TODO
+  Follows directly from $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$  since $\lrr(W_R) \subseteq \lrrg{\trust}(W_R)$.
+
 \item $\memSat[n']{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
-  TODO
+  By assumption.
+
 \item Theorem~\ref{thm:ftlr} holds for all $n'' < n'$:
   Follows from our induction hypothesis since $n' \le n$.
+
 \end{itemize}

### 2018-06-21 fully-abstract-comp/technical-report.tex
Remove obsolete commented out stuff.
Fix some open todos.

@@
   \label{lem:purePart-oplus}
-  For all $W_1,W_2$, we have that $\purePart{W_1\oplus W_2} = \purePart{W_1} \oplus \purePart{W_2}$.
-  Also, the latter is defined when the former is defined (but not necessarily vice versa).
+  For all $W_1,W_2$, we have that $\purePart{W_1\oplus W_2} = \purePart{W_1} \oplus \purePart{W_2} = \purePart{W_1} = \purePart{W_2}$.
+  Also, all worlds in the equations above are defined when $W_1\oplus W_2$ is defined (but not necessarily vice versa).
 \end{lemma}
@@
   \begin{itemize}
-  % \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
   \item $\npair{(w,\_)} \in \lrvg{\trusted}(W_w)$
-  % \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
-  % \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
-  % \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
-  % \item $\dom(\code) \subseteq \ta$
-  % \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
   \item $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$, with respect to $\code,\ta,\gsigrets,\gsigcloss$.
@@
     \npair{\left(\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
-      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lreg{\trust}(W)
+      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lre(W)
   \]
@@
   \item If $\npair{A} \in \execCond{W}$, then $\npair{A} \in \execCond{}(\purePart{W})$
+  \item If $\npair{A} \in \xReadCond{W}$, then $\npair{A} \in \xReadCond{}(\purePart{W} )$.
   \item If $\npair{(w_1,w_2)} \in \lrv(W)$ and ($\nonLinear{w_1}$ or

### 2018-06-21 fully-abstract-comp/ftlr-proof-cases.tex
Fix some TODOs.

@@
     \item $[\baddr,\eaddr] \subseteq \ta$
-    \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
+    \item ${((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
     \end{itemize}
@@
 
-By Lemma~\ref{lem:ftlr-internal-lemma}, it suffices to show that:
-\begin{itemize}
-\item $\Phi_S$ is reasonable up to $n'$ steps:
-  TODO
+By Lemma~\ref{lem:ftlr-internal-lemma}, taking $\trust = \trusted$ iff $[b,e] \subseteq \ta$ and $\trust = \untrusted$ otherwise, it suffices to show that:
+\begin{itemize}
+\item If $\trust =\trusted$ then $ \Phi_S$ is reasonable up to $n'$ steps.
+
+  We know by assumption that $c$ behaves reasonably up to $n$ steps.
+
+  By definition, it suffices to show that $\reg_S(r)$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$ for $r \nequal \pcreg$ and that $\ms_S$, $\ms_\stk$ and $\stk$ are all disjoint.
+
+  Take an $r \neq \pcreg$ and $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$.
+  By Lemma~\ref{lem:untrusted-source-values-are-reasonable}, it suffices to prove the following:
+  \begin{itemize}
+  \item $\npair{(w,\_)} \in \lrv(W_w)$: follows from $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$.
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$: by assumption.
+  \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined: By Lemma~\ref{lem:purePart-oplus}
+  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$: TODO
+  \item $\dom(\code) \subseteq \ta$: Follows from $\trust = \trusted$ by the choice of $\trust$.
+  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$: TODO
+  \end{itemize}
 
@@
 \item $\memSat[n']{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
+  By assumption.
+
+\item $W \oplus W_R \oplus W_M$ is defined:
   By assumption.

### 2018-06-21 fully-abstract-comp/ftlr-proof-cases.tex
FTLR register file safety case done.

--- 2018-06-19 fully-abstract-comp/ftlr-proof-cases.tex
@@
 \begin{proof}
-  TODO: update...
+  In the following we will use the following: $\reg_S = \Phi_S.\reg$, $\reg_T = \Phi_T.\reg$, $\ms_S = \Phi_S.\mem$, $\ms_\stk = \Phi_S.\ms_\stk$, and $\stk = \Phi_S.\stk$.
+
   By general induction on $n$, i.e. we can assume that the lemma already holds for $n' < n$.
@@
 
-  \lau{TODO: look through this case and find the place where the ''no seal cross boundary thing'' should have been used.}
+  \lau{TODO: look through this case and find the place where the ''no seal cross
+    boundary thing'' should have been used. (I have done this, and it does not
+    seem necessary (?))}
   In case \ref{case:ftlr:call}, we may assume $r_1 \neq r_2$\footnote{If the register contains a data capability, then the execution fails in the step after the jump.
@@
   \]
-  By assumption, we have $\npair[n']{(\reg_S(r_i),\reg_T(r_i))} \in \lrr(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
+  By assumption, we have $\npair[n']{(\reg_S(r_i),\reg_T(r_i))} \in \lrrg{\trust}(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
   We know the capabilities in $r_1$ and $r_2$ are sealed capabilities, and by
-  Lemma~\ref{lem:unique-h-sigma} and the definition of $\lrv$ we get
-  $\npair[n'-1]{\stpair[.]{c_2}{c_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$
+  Lemma~\ref{lem:unique-h-sigma} and the definition of $\lrvg{\trust}$ we get
+  $\npair[n'-1]{(c_2,c_2')} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$
   and w.l.o.g.\
@@
   We know $W_{R,1}' \future W_{R,1}$ as the revoked regions must have been $\spatial$ in $W_{R,q}$ (as they are owned by the part of the world assigned to the stack-register in the register-file relation).
-  The static region for the private stack is an extension.
+  The static region for the private stack is an extension of the old world.
 
@@
   \]
-  as desired. It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}.
+  as desired.
+  It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}, but first we note that we can deduce the following: From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\sigrets',\_,M(r),\gc}$.
+  Further by $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one $r \in S$ there is a $\mscode' \subseteq \mscode$, call this $r$ for $r_{\mscode'}$,\footnote{Follows from $H^\var{code}$ requiring $\dom(\mscode') \subseteq \ta$ and $\dom(\mscode) = \ta$ and finally that $\mscode$ is part of $\ms_S$.}
+  such that
+  \begin{equation}
+    \npair{(\mscode',\mscode')} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\_,\gsigrets,\gsigcloss)\label{eq:ftlr:rretd:Hcode}
+  \end{equation}
+  where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
+  That is $\mscode'$ contains the call we are considering.
+  This also entails
+  \begin{equation}
+    \label{eq:ftlr:code-comp-jud}
+    \sigrets',\sigcloss' \vdash_\var{code\_comp} \mscode'
+  \end{equation}
+from \ref{eq:ftlr:rretd:Hcode} we also get
+\begin{itemize}
+\item $\dom(\mscode') \subseteq \ta$
+
+  Otherwise we would have $\dom(\mscode') \# \ta$ which would contradict $\dom(\mscode') \subseteq \dom(\mscode) \subseteq\ta$
+
+\item $\sigrets' \subseteq \gsigrets$
+
+  Follows from the above.
+\end{itemize}
+  \lau{TODO: Make the above more precise and possibly put into a lemma (also needed in the next case).}
 
@@
   Case \ref{case:ftlr:scall:reg}: We need to split the ownership of $W_R'$.
-  From assumption $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
+  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
   We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership (i.e.\ $\purePart{W_{R,1}'}$).
@@
     \]
-    We know by Assumption~\ref{item:reg-ass}, we know
-    \[
-      \npair[n']{(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk},((\rw, \linear), \baddr_\stk ,\eaddr_\stk, \aaddr_\stk)} \in \lrrg{\trust}(W_{R,\stk})
-    \]
-    which by the $\stackReadCond{}$ gives $S \subseteq
-    \addressable{\linear,\pwfree}$ and $R : S \fun \powerset{\nats}$.
+    We know by Assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we know
+    \[
+      \npair[n']{(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk},((\rw, \linear), \baddr_\stk ,\eaddr_\stk, \aaddr_\stk)} \in \lrvg{\trust}(W_{R,\stk})
+    \]
+    which by the $\mathit{stackReadCondition}$ gives $S \subseteq \addressable{\linear,\pwfree}$ and $R : S \fun \powerset{\nats}$.
     We need to pick an $S'$ to argue
@@
     \[
-      \npair[n']([\baddr_\stk,\eaddr_\stk]) \in \stackReadCond{W_{R,\rstk}}
+      \npair[n']{[\baddr_\stk,\eaddr_\stk]} \in \stackReadCond{W_{R,\rstk}}
     \]
@@
     \]
-    Follows from the related register files assumption and Lemma~\ref{lem:lincons-lrv}.
+    Follows from the related register files assumption and Lemma~\ref{lem:lincons-lrv} and the fact that for sealed capabilities if they are in $\lrvg{\trusted}$, then they are in the $\lrv$ part.
 
@@
     where $W_{R,\rretd}'$ is the part of $W_R'$ with ownership over $r_\var{priv\_stk}$.
-    From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\sigrets',\_,M(r),\gc}$.
-    Further by $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one
-    $r \in S$ there is a $\mscode' \subseteq \mscode$\footnote{Follows from
-      $H^\var{code}$ requiring $\dom(\mscode') \subseteq \ta$ and $\dom(\mscode) =
-      \ta$ and finally that $\mscode$ is part of $\ms_S$.} such that
-    \begin{equation}
-      \npair[n']{(\mscode',\mscode')} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\_)\label{eq:ftlr:rretd:Hcode}
-    \end{equation}
-    where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
-    That is $\mscode'$ contains the call we are considering.
-    \lau{TODO: Make the above more precise and possibly put into a lemma (also needed in the next case).}
-    Use the $r$ in question as the witness.
+
+    Use the $r=r_{\mscode'}$ as the witness.
     The $\xReadCond{}$ gives us $\pwheap(r) \nequal \codereg{\_,\_,\mscode',\gc}$.
@@
       \]
-      First $\dom(\mscode') \subseteq \ta$ follows from \ref{eq:ftlr:rretd:Hcode} and the assumption $[\baddr,\eaddr] \ subseteq \ta$.
+      First, we already know $\dom(\mscode') \subseteq \ta$ and $\sigrets' \subseteq \gsigrets$.
       Now pick $r_\var{priv\_stk}$ as the witness.
       We immediately get $\dom(\ms_{\stk\_\priv,S}) = \dom(\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}) = [\aaddr_\stk,\eaddr_\stk]$.
-      Next, $\decInstr(\mscode'(\aaddr,\aaddr+\calllen-1)) = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ follows from $\callCond{\Phi_S,r_1,r_2,\aaddr}$.
+      Next, $\decInstr{\mscode'(\aaddr,\aaddr+\calllen-1)} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ follows from $\callCond{\Phi_S,r_1,r_2,\aaddr}$.
       Finally, $\ms_\code(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr}$ with $\sigma' = \sigma_\baddr + \offpc \in \sigrets'$ which follows from $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$ and the fact that the call is there.
@@
       $\isLinear{\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk)}}$\\
-      Trivial as both are linear.
+      Trivial, both are linear.
     \item
@@
       \]
-      Trivial as both configuration fails.
+      Trivial as both configurations fails.
     \end{enumerate}
@@
     where $W_{R,\rretc}'$ has the ownership of $W$.
-    Just as in the previous case, we know that for some region $r$ there exists $\mscode' \subseteq \mscode$ such that: $\pwheap(r) \nequal \codereg{\sigrets',\_,\mscode',\gc}$ where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
+    Just as in the previous case, we know that for some region $r$ there exists $\mscode' \subseteq \mscode$ such that: $\pwheap(r) \nequal \codereg{\sigrets',\_,\mscode',\gc}$ where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$ and $\sigrets' \subseteq \gsigrets$.
     It follows easily from the definition of $H_\sigma^\var{code}$ that
@@
     for $n'' < n'-1$.
-    Both capabilities are non-linear, so $\isLinear{\retptrc(\baddr,\eaddr,\aaddr + \calllen)}$ iff $\isLinear{((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)}$ is indeed the case.
+    Both capabilities are non-linear, so
+    \[
+      \isLinear{\retptrc(\baddr,\eaddr,\aaddr + \calllen)}\text{ iff } \isLinear{((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)}
+    \]
+    is indeed the case.
     Finally we need to show:
@@
     To this end let $W'' \future \purePart{W_{R,\rretc}'}$ and $W_o$ be given s.t.
-    $W'' \future \purePart{W_{R,\rretc}'} \oplus W_o$ is defined. Further, let
-    $\npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \;
-    \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \;
-    \xi^{-1}(W_o)$ be given and show
+    $W'' \oplus W_o$ is defined.
+    Further, let $\npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o)$ be given and show
     \[
@@
     In the first case, $\vsc_S'$ is a $\retptrc$ and $\vsc_T'$ is a capability with permission $\rx$.
-    In this case, $\xjumpResult{}{}{}$ will produce failed configurations which are trivially in the observation relation.
-    In the next case, it is required that $\sigma' \in \sigcloss'$, but this cannot be the case as \lau{TODO I think this case should follow from $H^\var{code}$ and the comp-code judgement, but it does not.
+    In this case, $\var{xjumpResult}$ will produce failed configurations which are trivially in the observation relation.
+    In the next case, it is required that $\sigma' \in \sigcloss'$, but this
+    cannot be the case as $\sigma' \in \sigrets'$ and we have $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$.
+\lau{TODO I think this case should follow from $H^\var{code}$ and the comp-code judgement, but it does not.
       Specifically, the comp-code judgement (for now) does not require $\sigrets'$ and $\sigcloss'$ to be disjoint.}
@@
       \begin{array}{l}
-        \exists r \in \addressable{\linear,\pwpriv[W_o]} \ldotp \pwpriv[W_o](r).H \nequal (\stareg[(\ms_{\priv,S}^{(3)},\ms_{\priv,T}^{(3)}),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
-        \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr_\stk',\eaddr_\stk'] \tand\\
+        \exists r \in \addressable{\linear,\pwpriv[W_o]} \ldotp \pwpriv[W_o](r).H \nequal (\stareg[(\ms_{\priv,S}^{(3)},\ms_{\priv,T}^{(3)}),\gc]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \quad \dom(\ms_{\priv,S}^{(3)}) = \dom(\ms_{\priv,T}^{(3)}) = [\baddr_\stk',\eaddr_\stk'] \tand\\
         \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
@@
 
-    We know that the two register-files are related which in particular means that
-    the values in register $\rstk$ are related. Now consider the following cases:
-    \lau{TODO: Perhaps make a lemma that says if source fail and target step to
-      fail, then the two configurations are in $\lro$.}
-    \begin{itemize}
-    \item $\Phi_S(\rstk) \neq \stkptr{\_,\_,\_,\_}$\\
+    We know that the two register-files are related which in particular means that the values in register $\rstk$ are related.
+    Now consider the following cases: \lau{TODO: Perhaps make a lemma that says if source fail and target step to fail, then the two configurations are in $\lro$.}
+    \begin{itemize}
+    \item $\reg_S^{(3)}(\rstk) \neq \stkptr{\_,\_,\_,\_}$\\
       In this case due to $\reg[3]_S$ being related to $\reg[3]_T$, there are three cases we need to consider.
@@
       This means that the execution will fail that $\stkb$ check.
-    \item $\Phi_S(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\_,\_}$ and $\baddr_\var{ret\_\stk} \neq \stkb$\\
+    \item $\reg_S^{(3)}(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\_,\_}$ and $\baddr_\var{ret\_\stk} \neq \stkb$\\
       Here the source side will fail the xjmp as the base address is not $\stkb$.
       Similarly on the target side, the return code will fail the $\stkb$ check.
-    \item $\Phi_S(\rstk) = \stkptr{\_,\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and either $\eaddr_\var{ret\_\stk} +1 \neq \baddr_\stk' $ or $\perm_\var{ret\_\stk} \neq \rw$ or
+    \item $\reg_S^{(3)}(\rstk) = \stkptr{\_,\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and either $\eaddr_\var{ret\_\stk} +1 \neq \baddr_\stk' $ or $\perm_\var{ret\_\stk} \neq \rw$ or
       $\baddr_\var{ret\_\stk} > \eaddr_\var{ret\_\stk}$.\\
@@
       On the target side, the splice will fail as either the two capabilities being spliced don't line up, the permissions don't match, or the range of authority is empty, respectively.
-    \item $\Phi_S(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and $\eaddr_\var{ret\_\stk} +1 = \baddr_\stk' $ and $\perm_\var{ret\_\stk} = \rw$ and $\baddr_\var{ret\_\stk} \leq \eaddr_\var{ret\_\stk}$.
+    \item $\reg_S^{(3)}(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and $\eaddr_\var{ret\_\stk} +1 = \baddr_\stk' $ and $\perm_\var{ret\_\stk} = \rw$ and $\baddr_\var{ret\_\stk} \leq \eaddr_\var{ret\_\stk}$.
       \\
@@
         \]
-        Here we need to use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$.
-        TODO: continue
-        \lau{TODO: the reasonability requirements of the FTLR seems to be an issue here.}
-        \dominique{should now be fine?}
+        For fresh $r_{b_\stk'}\dots r_{e_\stk'}$ and $W'''$ defined as
+        \[
+          W''' = W''[\priv.r' \mapsto \revoked,\free.r_{b_\stk'}\dots r_{e_\stk'}\mapsto \stdreg{\{b_\stk'\},\gc}{\spa} \dots \stdreg{\{e_\stk'\},\gc}{\spa}]
+        \]
+        and $W'''_R$ with the ownership of $W''_R$ as well as for the regions $r_{b_\stk'}\dots r_{e_\stk'}$,
+        and $W'''_M$ with the ownership of $W''_M$,
+        we show the following:
+        \begin{enumproof}
+        \item $\npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrv(W'''_R)$
+        \item $\memSat[n''']{\Phi_S^{(4)}.\mem,\Phi_S^{(4)}.\stk,\Phi_S^{(4)}.\ms_\stk,\Phi_T^{(5)}.\mem}{W'''_M}$ 
+        \item $W''' \oplus W'''_R \oplus W'''_M$ defined
+        \end{enumproof}
+        Assuming the above, we use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$ to get
+        \[
+          \npair[n''']{(((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen))} \in \lreg[\trusted](W''')
+        \]
+        Using this along with the register-file safety and memory satisfaction,
+        we get
+        \[
+          \npair[n''']{(\Phi^{(4)}_S,\Phi^{(5)}_T)} \in \lro
+        \]
+        as desired. We need to show the three things we skipped:
+
+        Show:
+        \[
+          \npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrrtrusted(W'''_R)
+        \]
+        Using Lemma~\ref{lem:untrusted-subset-trust}, it suffices to show $\npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrrtrusted(W'''_R)$.
+        We will split the world like in $\npair[n''']{(\reg_S^{(3)},\reg_T^{(3)})} \in \lrr(W)$, but where $r_\stk$ also get the ownership of $r_{b_\stk'}\dots r_{e_\stk'}$.
+        We need to show the following
+        \begin{itemize}
+        \item Case $\rdata$, $\rtmp{1}$, $\rtmp{2}$:
+
+          Trivial.
+        \item Case $r \not\in \{\rdata,\rtmp{1},\rtmp{2},\pcreg,\rstk\}$:
+
+          We have $\Phi^{(4)}_S.\reg(r) = \reg^{(3)}_S(r)$ and $\Phi^{(5)}_T.\reg(r) = \reg^{(3)}_T(r)$.
+          We already know: $\npair[n''']{(\reg^{(3)}_S(r),\reg^{(3)}_T(r))} \in \lrr(W''_{R,r})$, so by Lemma~\ref{lem:monotonicity}, we get
+          \[
+            \npair[n''']{(\reg^{(3)}_S(r),\reg^{(3)}_T(r))} \in \lrr(W'''_{R,r})
+          \]
+          as desired.
+        \item Case $\rstk$:
+
+          For this case, we need to show
+          \[
+            \npair[n''']{[\stkb,e_\stk']} \in \stackReadCond{W'''_{R,\rstk}}
+          \]
+          and
+          \[
+            \npair[n''']{[\stkb,e_\stk']} \in \stackWriteCond{W'''_{R,\rstk}}
+          \]
+          For $W'''_{R,\rstk}$ that owns the same as $W''_{R,\rstk}$ as well as $r_{b_\stk'},\dots,r_{e_\stk'}$.
+
+          For the first part, we use $\npair[n''']{(\reg_S^{(3)}(\rstk),\reg_T^{(3)}(\rstk))} \in \lrv(W_{R,\rstk}'')$ and the fact that the stack capability must have $\rw$ permission from which it follows that
+          \[
+            \npair[n''']{[\stkb,e_\var{ret\_stk}]} \in \stackReadCond{W''_{R,\rstk}}
+          \]
+          which gives us $S_\free \subseteq \addressable{\linear,\pwfree}$ and $R_\free : S_\free \fun \powerset{\nats}$ such that
+          \begin{itemize}
+          \item $\forall r \in S_\free \ldotp |R(r)| = 1$
+          \item $\biguplus_{r \in S_\free} R(r) \subseteq [\stkb,e_\var{ret\_stk}]$
+          \item $\forall r \in S_\free \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\spao}$
+          \end{itemize}
+          Now pick $S_\var{read} = S_\free \cup \{r_{b_\stk'},\dots,r_{e_\stk'}\}$ and
+          \[
+            R_\var{read}(r) =
+            \begin{cases}
+              R_\var{read}(r) & r \in S_\var{read} \\
+              \{a\} & r \in \{r_{b_\stk'},\dots,r_{e_\stk'}\} \wedge r = r_a
+            \end{cases}
+          \]
+          It is clearly the case that $\forall r \in S_\var{read} \ldotp |R_\var{read}(r)| = 1$ and $\biguplus_{r \in S_\var{read}} R(r) \subseteq [\stkb,e_\var{ret\_stk}]$.
+          For $\forall r \in S_\var{free} \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\spao}$ it follows from $\forall r \in S_\free \ldotp \pwfree(r) \nsupeq \stdreg{R(r),\gc}{\spao}$ and $\pwfree(r) = \stdreg{R(r),\gc}{\spao}$ for $r \in \{r_{b_\stk'},\dots,r_{e_\stk'}\}$.
+
+          The $\var{stackReadCondition}$ is showed in the same way, but we also use Lemma~\ref{lem:stdreg-singleton-addr-strat} to argue that the new regions are address-stratified.
+        \end{itemize}
+
+        Show:
+        \begin{equation}
+          \memSat[n''']{\ms^{(3)}_S,\Phi_S^{(4)}.\stk,\ms_\stk^{(3)} \uplus \ms_{\priv,S}^{(3)} ,\ms^{(3)}_T}{W'''_M}\label{eq:ftlr:case:last-mem-sat}
+        \end{equation}
+
+        From the assumption $\memSat[n''']{\ms^{(3)}_S,\ms^{(3)}_\stk, \stk^{(3)}, \ms^{(3)}_T}{W_M''}$, we know
+        \begin{itemize}
+        \item $\stk = (\addr + \calllen,\ms_{\priv,S}^{(3)})::(\opc_1,\ms_1):: \dots :: (\opc_m,\ms_m)$
+        \item $\ms^{(3)}_S \uplus \ms^{(3)}_\stk \uplus \ms_{\priv,S}^{(3)} \uplus \ms_1 \uplus \dots \uplus \ms_m  \text{ is defined}$
+        \item $W''_M = W_{\var{stack}}'' \oplus W_{\var{free\_stack}}'' \oplus W_{\var{heap}}''$
+        \item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
+          \begin{itemize}
+          \item $\ms^{(3)}_S = \ms_{f,S} \uplus \ms_S'$
+          \item $\ms^{(3)}_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
+          \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}''}$
+          \item $\memSatFStack{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$
+          \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}}'')$
+          \end{itemize}
+        \end{itemize}
+        We will pick the same things to show \ref{eq:ftlr:case:last-mem-sat} with a few changes.
+        We have to show
+        \begin{itemize}
+        \item $\Phi_S^{(4)}.\stk = (\opc_1,\ms_1):: \dots :: (\opc_m,\ms_m)$
+
+          By the memory satisfaction assumption and the change to the stack.
+        \item $\ms^{(3)}_S \uplus \ms^{(3)}_\stk \uplus \ms_{\priv,S}^{(3)} \uplus \ms_1 \uplus \dots \uplus \ms_m \text{ is defined}$
+
+          By the memory satisfaction assumption.
+        \item $W'''_M = W_{\var{stack}}''' \oplus W_{\var{free\_stack}}''' \oplus W_{\var{heap}}'''$
+
+          Define the new worlds to have the ownership of their $W_M''$ counterparts except $W_{\var{stack}}'''$ does not take ownership of the regions used for the safety of addresses $b_\stk',\dots,e_\stk'$.
+          This ownership goes to $W_{\var{free\_stack}}'''$ instead.
+        \item We need to pick partitions of $\ms^{(3)}_T$ and a frame for $\ms^{(3)}_S$.
+          We pick the same as we get from the memory satisfaction assumption except we pick the free stack partition of the target memory to be $\ms_\var{T,free\_stack} \uplus \ms_{\priv,S}^{(3)}$ and the stack partition to be $\ms_\var{T,stack}\setminus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}$
+        \item $\ms^{(3)}_S = \ms_{f,S} \uplus \ms_S'$
+
+          By assumption.
+        \item $\ms^{(3)}_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
+
+          By assumption and the fact that the only change is that we moved part of the stack to the free stack.
+        \item $\memSatStack{(\opc_1,\ms_1):: \dots :: (\opc_m,\ms_m),\ms_\var{T,stack}\setminus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}}{W_{\var{stack}}'''}$
+
+          Follows easily from the private stack satisfaction assumption.
+          The distribution functions from the assumption are simply limited to forget about the now revoked region and extend the world partition to be on $W_M'''$, but with the same ownership as in the one we had in the assumption.
+          All the new partitions are future worlds of the old ones as none of them owned the revoked region (it was owned by $W_R''$).
+        \item $\memSatFStack{\ms^{(3)}_\stk \uplus \ms_{\priv,S}^{(3)},\ms_\var{T,free\_stack} \uplus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}}{W_{\var{free\_stack}}'''}$
+
+          From the $\memSatFStack{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$ assumption we get $R_\ms : \dom(\activeReg{W_{\var{free\_stack}}''}) \fun \MemSeg \times \MemSeg$ and $R_W : \dom(\activeReg{W_{\var{free\_stack}}''}.\free) \fun \Worlds$ for which
+          \begin{itemize}
+          \item $\ms^{(3)}_\stk = \biguplus_{r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free})} \pi_1(R_\ms(r))$
+          \item $\ms_\var{T,free\_stack} = \biguplus_{r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free})} \pi_2(R_\ms(r))$
+          \item $\stkb \in \dom(\ms^{(3)}_\stk)$ and $\stkb \in \dom(\ms_\var{T,free\_stack})$
+          \item $W_{\var{free\_stack}}'' \bigoplus_{r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free})} R_W(r)$
+          \item $\forall r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free})\ldotp \npair[n''']{R_\ms(r)} \in W_{\var{free\_stack}}''.\free(r).H \; \xi^{-1}(R_W(r))$
+          \end{itemize}
+          Now pick
+          \[
+            R_\ms'(r) =
+            \begin{cases}
+              (\ms_{\priv,S}^{(3)}|_{\{a'\}} , \ms_\var{T,stack}|_{\{a'\}}) & r_{a'} \in \{r_{b_\stk'}\dots r_{e_\stk'}\}\\
+              R_\ms(r) & \totherwise
+            \end{cases}
+          \]
+          and
+          \[
+            R_W'(r) =
+            \begin{cases}
+              W'''_{a'} & r_{a'} \in \{r_{b_\stk'}\dots r_{e_\stk'}\} \\
+              R_W(r)[\priv.r' \mapsto \revoked,\free.r_{b_\stk'}\dots r_{e_\stk'}\mapsto \stdreg{\{b_\stk'\},\gc}{\spa} \dots \stdreg{\{e_\stk'\},\gc}{\spa}]& \totherwise
+            \end{cases}
+          \]
+          for $W'''_{a'}$ constructed as follows: $W_{\var{stack}}''$ is the part of the world given to the stack judgement in the assumption.
+          This world is split into a number of parts to satisfy the memory interpretation of each frame.
+          Say $W_{top,\var{stack}}''$ is used for the top stack frame in the assumption.
+          The top stack frame is governed by a static region, so by definition it is split into parts that satisfy each of the addresses.
+          That is for $a' \in \{b_\stk' \dots e_\stk'\}$, $W_{a'}''$ is the part that makes the value in memory satisfy the value relation.
+          Now let $W_{a'}'''$ be $W'''$ but with the ownership of $W_{a'}''$
+
+          It is easy to see that the $R_\ms'$ constructs the two memories, and
+          from the assumption, we also get that the base stack address is in there.
+
+          It remains to show
+          \[
+            \forall r \in \dom(\activeReg{W_{\var{free\_stack}}'''.\free}) \in \npair[n''']{R_\ms(r)} \in W_{\var{free\_stack}}'''.\free(r).H \; \xi^{-1}(R_W(r))
+          \]
+          for $r \in \dom(\activeReg{W_{\var{free\_stack}}'''.\free}) \setminus \{r_{b_\stk'}\dots r_{e_\stk'}\}$, it follows from monotonicity of the $H$ (memory interpretation) function.
+          For $r_{a'} \in \{r_{b_\stk'}\dots r_{e_\stk'}\}$, we need to show
+          \[
+            \npair[n''']{(\ms_{\priv,S}^{(3)}(a') , \ms_\var{T,stack}|(a'))} \in \stdreg{\{a\},\gc}{\spa}.H (\xi^{-1}R_W(r_{a'}))
+          \]
+          which amounts to showing $\dom(\ms_{\priv,S}^{(3)}|_{\{a'\}}) = \dom(\ms_\var{T,stack}|_{\{a'\}}) \{a'\}$, which is the case, and
+          \[
+            \npair[n''']{(\ms_{\priv,S}^{(3)}(a') , \ms_\var{T,stack}|(a'))} \in \lrv (W_{a'}''')
+          \]
+          Which follows from Lemma~\ref{lem:monotonicity} and the fact that we have a memory satisfaction assumption in which $\ms_{\priv,S}^{(3)}$ is governed by a standard static region.
+
+        \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}}''')$
+
+          Follows by region interpretation monotonicity and the heap satisfaction assumption.
+        \end{itemize}
+
+        Argue:
+        \[
+          W''' \oplus W'''_R \oplus W'''_M \text{ is defined.}
+        \]
+        This follows from the assumption $W'' \oplus W''_R \oplus W''_M$ and the fact that each of the new worlds is constructed from one of the past worlds and only one of them claims the ownership of the new regions.
       \end{itemize}
@@
 
+\begin{lemma}
+  \label{lem:stack-sat-unchanged-future}
+  If $\pwfree = \pwfree[W']$ and $W' \future W$ and
+  $\memSatFStack{\ms_\stk,\ms_T}{W}$, then $\memSatFStack{\ms_\stk,\ms_T}{W'}$.
+\end{lemma}
+
+\begin{lemma}
+  \label{lem:stdreg-singleton-addr-strat}
+  $\stdreg{\{*\},\gc}{v}$ is address-stratified.
+\end{lemma}
+\begin{proof}
+  Trivial.
+\end{proof}
+
 \begin{lemma}[Unique return seals]

### 2018-06-21 fully-abstract-comp/technical-report.tex
Fix some TODOs.
FTLR register file safety case done.

--- 2018-06-20 fully-abstract-comp/technical-report.tex
@@
       \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
+      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)} \wedge\\
       \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
@@
 \]
+
+\begin{lemma}[Untrusted is trusted]
+  \label{lem:untrusted-subset-trust}
+  \begin{itemize}
+  \item $\lrvtrusted(W) \subseteq \lrv(W)$
+  \item $\lrrtrusted(W) \subseteq \lrr(W)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
 
@@
     \begin{array}{l}
-    \dom(\code) = [\aaddr,\baddr] \wedge \\
-      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \trust = \trusted) \vee \\
-      \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
+      ((\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \trust = \trusted) \vee \\
+      \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)) \wedge \\
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
@@
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to
-      $n'-i$ steps
+    \item $\Phi'(\pcreg) + \calllen$ behaves reasonably up to $n'-i$ steps
     \end{itemize}
@@
   \label{lem:codereg-Hsigma-not-lrv}
-  If $\dom(\code) \subseteq \ta$,  $\sigrets \subseteq \gsigrets$, $\sigcloss \subseteq \gsigcloss$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)$.
+  If $\dom(\code) \subseteq \ta$,  $\sigrets \subseteq \gsigrets$, $\sigcloss \subseteq \gsigcloss$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)$.
 \end{lemma}
@@
 
-    In this case, we have that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but we can show that it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$ by taking a $\hat{W}$ such that the following hold:
+    In this case, we have that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but we can show that it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~\hat{W}$ by taking a $\hat{W}$ such that the following hold:
     \begin{itemize}
@@
   \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
-  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
+  \item $\pwheap[W_w](r_\code) \nequal \codereg{\gc}$
   \item $\dom(\code) \subseteq \ta$
@@
 
-    From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
+    From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
     However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
@@
   \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
-  \item $\pwheap[W_w](r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
+  \item $\pwheap[W_w](r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
   \item $\dom(\code) \subseteq \ta$
@@
     \item $r \in \dom(\pwheap)$
-    \item $\pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
+    \item $\pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}$
     \item $\dom(\code) \subseteq \ta$ 
@@
   Follows easily from the definitions.
-  In the case for $\execCond$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
+  In the case for $\execCond{}$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
   It suffices construct new $W_a'$ by taking $W_a$ but adding $\oplus_{a\in(A\setminus A')}$ to an $W_a(a)$ for an arbitrary $a \in A'$ to make this work.

### 2018-06-22 fully-abstract-comp/technical-report.tex
Proper indices in reasonability definitions.

--- 2018-06-21 fully-abstract-comp/technical-report.tex
@@
       \begin{itemize}
-      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n$ steps.
-      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
+      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n' - i$ steps.
+      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n' - i$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
       \end{itemize}
@@
     \end{itemize}
-    We have that $\Phi$ is reasonable.
+    We have that $\Phi$ is reasonable up to $n$ steps.
   \end{definition}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Modify step-indexing in the reasonability definition to fix Lau's problem in the call case.
Remove the i > 0 requirements elsewhere in the reasonability definitions too, as for call.

@@
       for some $\src{r_1}$ and $\src{r_2}$
-    \item $0 < i \le n'$
+    \item $i \le n'$
     \end{itemize}
@@
     \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to
-      $n'-i$ steps
+      $n'-i - 1$ steps
     \end{itemize}
@@
     \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item i > 0
     \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
@@
       \begin{itemize}
-      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n' - i$ steps.
-      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n' - i$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
+      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n' - i - 1$ steps.
+      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n' - i - 1$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
       \end{itemize}
@@
     \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
-    \item $0 < i \le n'$
-    \end{itemize}
-    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n'-i$ steps
+    \item $i \le n'$
+    \end{itemize}
+    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n'-i -1$ steps
   \item \emph{Don't leak private stuff...} If
@@
     \item $\src{\Phi''}$ is not in $\ta$
-    \item $0 < i \le n'$
+    \item $i \le n'$
     \end{itemize}

### 2018-06-22 fully-abstract-comp/ftlr-proof-cases.tex

@@
 
-  By general induction on $n$, i.e. we can assume that the lemma already holds for $n' < n$.
+  By complete induction on $n$, i.e. we can assume that the lemma already holds for $n' < n$.
 
@@
     \item $[\baddr,\eaddr] \subseteq \ta$
-    \item $\trust = \trusted$
-    \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
+    \item ${((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
     \end{itemize}
@@
 
-By Lemma~\ref{lem:ftlr-internal-lemma}, it suffices to show that:
-\begin{itemize}
-\item $\Phi_S$ is reasonable up to $n'$ steps:
-  TODO
-\item $\trust = \trusted \vee \npair[n']{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$:
-  TODO
+By Lemma~\ref{lem:ftlr-internal-lemma}, taking $\trust = \trusted$ iff $[b,e] \subseteq \ta$ and $\trust = \untrusted$ otherwise, it suffices to show that:
+\begin{itemize}
+\item If $\trust =\trusted$ then $ \Phi_S$ is reasonable up to $n'$ steps.
+
+  We know by assumption that $c$ behaves reasonably up to $n$ steps.
+
+  By definition, it suffices to show that $\reg_S(r)$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$ for $r \nequal \pcreg$ and that $\ms_S$, $\ms_\stk$ and $\stk$ are all disjoint.
+
+  Take an $r \neq \pcreg$ and $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$.
+  By Lemma~\ref{lem:untrusted-source-values-are-reasonable}, it suffices to prove the following:
+  \begin{itemize}
+  \item $\npair{(w,\_)} \in \lrv(W_w)$: follows from $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$.
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$: by assumption.
+  \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined: By Lemma~\ref{lem:purePart-oplus}
+  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$: TODO
+  \item $\dom(\code) \subseteq \ta$: Follows from $\trust = \trusted$ by the choice of $\trust$.
+  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$: TODO
+  \end{itemize}
+
+\item $\trust = \trusted \vee \npair[n']{[\baddr,\eaddr]} \in \readCond{\lin,W}$:
+  If $\trust =\ untrusted$ then we know that $[b,e] \mathrel{\#} \ta$, so that $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ follows by Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
+
 \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\aaddr)$:
-  TODO
-\item $\npair[n']{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
-  TODO
+  We know that $\Phi_S(\pcreg) = \Phi_T(\pcreg) = c = ((\rx,\lin),\baddr,\eaddr,\aaddr)$.
+
+\item $\npair[n']{[\baddr,\eaddr]} \in \xReadCond{W}$:
+  Follows from $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ using Lemma~\ref{lem:downwards-closed}.
+
 \item $\npair[n']{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
-  TODO
+  Follows directly from $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_R)$  since $\lrr(W_R) \subseteq \lrrg{\trust}(W_R)$.
+
 \item $\memSat[n']{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
-  TODO
+  By assumption.
+
+\item $W \oplus W_R \oplus W_M$ is defined:
+  By assumption.
+
 \item Theorem~\ref{thm:ftlr} holds for all $n'' < n'$:
   Follows from our induction hypothesis since $n' \le n$.
+
 \end{itemize}

### 2018-06-22 fully-abstract-comp/technical-report.tex

--- 2018-06-21 fully-abstract-comp/technical-report.tex
@@
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ behaves reasonably up to $n'-i$ steps
+    \item $\Phi'(\pcreg) + \calllen$ behaves reasonably up to $n'-i - 1$ steps
     \end{itemize}
@@
       \begin{itemize}
-      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n$ steps.
-      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
+      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n' - i - 1$ steps.
+      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n' - i - 1$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
       \end{itemize}

### 2018-06-22 fully-abstract-comp/technical-report.tex
added a possibly i <= n' in reasonability for configurations
typo

@@
     \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
+    \item $i \le n'$
     \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
@@
   \label{def:points-to-call}
-  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{o}{r_1}{r_2}$ in $A$ iff
+  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{\offpc,\offsigma}{r_1}{r_2}$ in $A$ iff
   \[

### 2018-06-22 fully-abstract-comp/technical-report.tex
Comment out obsolete notes.

--- 2018-06-22 fully-abstract-comp/technical-report.tex
@@
 %\lau{15-12-2017: Removed linearity from seal - I could not come up with a single good reason to have it as we have them here.}
-\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
-\lau{28-09-2017: Why would you like this?
+% \dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
+% \lau{28-09-2017: Why would you like this?
 If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail.
 We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
-\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
-\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part.
+% \dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
+% \lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part.
 Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}.
@@
 Makes sense?}
+% \dominique{22-6-2018: yes}
 The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
@@
 The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc and 2) caller's private stack.
-\lau{18-12-2017: Stack frame used to contain the address of the old stack pointer. This should be inferable from other information, so it was removed.}
+% \lau{18-12-2017: Stack frame used to contain the address of the old stack pointer. This should be inferable from other information, so it was removed.}
 
@@
 \noindent\textbf{call}\\
-\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
-\lau{19-09-2017: In which cases do we have undefined here?}
-\lau{25-10-2017: I have forgotten why we save $e_\stk$ on the stack and whether this is still necessary (it at least prevent the empty case, but I do not recall if it is necessary for other reasons).}
-\lau{20-12-2017: At some point, I added an offset annotation $o$ to all calls.
+% \lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
+% \lau{19-09-2017: In which cases do we have undefined here?}
+% \lau{25-10-2017: I have forgotten why we save $e_\stk$ on the stack and whether this is still necessary (it at least prevent the empty case, but I do not recall if it is necessary for other reasons).}
+% \lau{20-12-2017: At some point, I added an offset annotation $o$ to all calls.
 I believe this removes the need for a function that annotate seals.
 It will still have to be checked that all calls in a program uses different offsets, but I suspect this can be done as some assumption in what we want to prove.}
-\lau{20-12-2017: Until this day, we had a $\sealAss{}$ function: This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack.
+% \lau{20-12-2017: Until this day, we had a $\sealAss{}$ function: This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack.
 We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}}
@@
 \end{multline*}
-\lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$.
+% \lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$.
 There is, however, no check in the implementation of call that ensures this.
@@
 And if we do not check it as part of the call implementation, do we then want to match the call implementation better?}
-\dominique{13-3-2018: re the above note: the above should be mitigated by Definition~\ref{def:check-stack-addr-before-call}.
+% \dominique{13-3-2018: re the above note: the above should be mitigated by Definition~\ref{def:check-stack-addr-before-call}.
 Adding a check is in principle not necessary, since the check would only protect trusted code from itself, rather than the adversary.}
-\dominique{13-3-2018: where does $o$ come from above in the calculation of $\sigma_?$?}
+% \dominique{13-3-2018: where does $o$ come from above in the calculation of $\sigma_?$?}
 

### 2018-06-22 fully-abstract-comp/technical-report.tex
Some minor cleanups.
Remove some more obsolete comments.
We're not using lre-trusted anymore, so remove for simplicity.
Latex cleanup.
Remove obsolete comments.

@@
 
-% For both $\privft$ and $\pubft$ we have
-% \begin{mathpar}
-%   \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
-%             { W' \future W }
-% \end{mathpar}
-%
 Future regions allow $\spatial$ regions to become $\revoked$.
@@
 \end{mathpar}
-%
-% Public future regions are allowed to transition according to the public part of their transition system.
-% \begin{mathpar}
-%     \inferrule{ }
-%             { \revoked \pubft \revoked }
-%             \and
-%             \inferrule{}{(\spatialo,H) \privft (\spatial,H)}
-%   %           \and
-%   % \inferrule{ (s,s') \in \phi_\pub}
-%   %           { (\spatialo,s',\phi_\pub,\phi,H) \pubft (\spatial,s,\phi_\pub,\phi,H)  }
-%   %           \and
-%   % \inferrule{ (s,s') \in \phi_\pub}
-%   %           { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
-% \end{mathpar}
 
@@
 
-\begin{lemma}[Adding additional memory is the future]
+\begin{lemma}[Adding memory is the future]
   \label{lem:uplus-future}
@@
   Follows easily from the definitions.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:comp-world-pure-comp-world}
-  If $W_1 \oplus W_2$ is defined, then $\purePart{W_1} \oplus \purePart{W_2}$ is defined
-\end{lemma}
-\begin{proof}
-  Follows from Lemma~\ref{lem:purePart-oplus}.
 \end{proof}
@@
 \]
-% }
 
@@
 We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints imposed by the world are satisfied.
-% \lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
-% \dominique{4-12-2017: Regarding the question whether $\opc$ should be mentioned in the memory satisfaction relation.
-%   The relevant intuitive aspect of the proof here, I think, is that on a return, the source semantics restores opc from the copy kept in the stack, while the target semantics restores it from the copy in the code part of the return pointer.
-%   We will need to prove (somewhere) that these two correspond, which means we will need to track of the invariant that guarantees this somewhere (perhaps here?).
-%   I suspect the relevant invariants are (1) that any executable pointer signed with $\sealAss{a}$ must point to $a + \calllen$ and (2) that anything related to a $\rretc$ must be executable and sealed.
-%   Perhaps this invariant is easiest to track in (1) the general requirement on the $H_\sigma$ and (2) in the value relation.
-%   This suggests nothing should be required about opc here?
-%   To be confirmed in the proof of the fundamental theorem.
-% }
 
@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\untrusted}(\{\rdata\}) (W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
@@
     \right\}\\
-  \lreg{\trust}(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
+  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
     \begin{array}{l}
@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust} (W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
@@
 We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
-% DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
-% Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
-%\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
-% Lau 19-01-2018: Added rdata as default.
 \lau{RESOLVED check what conditions need to have a ``square'' i.e. need approximation}
@@
 \lau{RESOLVED: Make this condition require a set of adjacent regions.}
-\lau{TODO: allow code regions to be merged into one. Allow for (... forgot to write the rest, but probably related to splicing regions.) }
 
@@
   \begin{itemize}
-  \item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\untrusted}(\{\rdata\}) (W_\lrrs )$
+  \item $\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs )$
   \item $\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm}$

### 2018-06-22 fully-abstract-comp/technical-report.tex
Remove obsolete definition.
Remove obsolete comment.
More minor cleanups.

@@
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
-% \dominique{Why $n-1$ above?
-There is already a quantification over $n' <n$ in the memory satisfaction relation and another $n-1$ is used in $\stdreg{[\baddr',\eaddr'],\gc}{\pur}$.
-Also: readCond doesn't have the $n-1$.
-Update: the memory satisfaction was updated by Lau to remove the quantification over $n' < n$ and I dropped the $n-1$ here.}
 \[
@@
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
-
-% \[
-%   \execCond{\lin,W} = \left\{ \npair{(S_\perm,\baddr,\eaddr)} \middle|
-%     \begin{array}{l}
-%       \forall n' < n, W' \privft W, \perm \in S_\perm, \aaddr \in [\baddr,\eaddr] \ldotp\\
-%       \quad \forall W'', c_{d,S},c_{d,T} \ldotp \npair[n']{\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lrv(W'') \wedge \\
-%       \qquad \npair[n']{[\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}]} \in \lrp(W' \oplus W'') \Rightarrow \\
-%       \quad \qquad \npair[n']{\stpair[.]{((\perm,\lin),\baddr,\eaddr,\aaddr)}{((\perm,\lin),\baddr,\eaddr,\aaddr)},\stpair[.]{c_{d,S}}{c_{d,T}}} \in \lre(W' \oplus W'')
-%     \end{array}
-%     \right\}
-% \]
-
-% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
-It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
+\dominique{22-6-2018: address-stratified is useless for length-one regions.}
+
 Note: this new version of execCond uses the expression relation for regular jumps since it expresses the validity of jumping to an address pointed to by an executable capability. 
@@
 \]
-\lau{01-11-2017: RESOLVED look into what kind of world separation should happen here.}
 
@@
 \]
-
-% \[
-%   \staureg{v} = (v,*,=,=,H^{\mathrm{sta},\lrv}_{\stpair{\ms}{\ms}})
-% \]
-% \[
-%   H^{\mathrm{sta},\lrv}_{\stpair{\ms}{\ms}} = 
-%     \left\{ \npair{\stpair{\ms}{\ms}} \middle|
-%       \begin{array}{l}
-%         \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\src{\ms_S})} S(\aaddr) \wedge\\        
-%         \quad \forall \aaddr \in \dom(\ms) \ldotp \npair[n-1]{\src{\ms_S}(\aaddr),\ms_T(\aaddr)} \in \lrv(S(a))
-%       \end{array}
-% \right\}
-% \]
 
@@
 
-% \subsection{World specifies component}
-
-% \begin{definition}[World specifies component]
-%   We define that a world $W$ specifies the presence of a component with global constants $\gc$
-%   \begin{equation*}
-%     \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
-%   \end{equation*}
-%   (written as $W \vdash_\gc \var{comp}_0$)
-%   iff there exist $r_1 \cdots r_n \in \dom(\pwheap)$ such that for all $i$,
-%   \begin{itemize}
-%   \item 
-%     \begin{equation*}
-%       \vdash (\code_i,\msdata[i],\overline{c_{\mathrm{import},i}},\overline{c_{\mathrm{export},i}},\sigrets_i,\sigcloss_i,A_{\linear,i})
-%     \end{equation*}
-%   \item $\mscode = \biguplus_i \code_i$
-%   \item $\msdata = \biguplus_i \msdata[i]$
-%   \item $\overline{c_{\mathrm{export}}} = \biguplus_i \overline{c_{\mathrm{export},i}}$
-%   \item $\overline{c_{\mathrm{import}}} = \biguplus_i \overline{c_{\mathrm{import},i}} \setminus \overline{c_{\mathrm{export}}}$
-%   \item $\sigrets = \biguplus \sigrets_i$
-%   \item $\sigcloss = \biguplus \sigcloss_i$
-%   \item $A_\linear = \biguplus A_{\linear,i}$
-%   \item $\pwheap(r_i) = \codereg[\mathrm{code}_i,\square]{\sigrets_i,\sigcloss_i,\code_i,\gc}$
-%   \item For all $i$, one of the following holds:
-%     \begin{itemize}
-%     \item $\dom(\code_i) \subseteq \ta$
-%     \item $\dom(\code_i) \mathrel{\#} \ta$
-%     \end{itemize}
-%   \end{itemize}
-%   and there exist $r_{\mathrm{addr}} : \dom(\msdata) \mapsto \dom(\pwheap)$ such that
-%   \begin{itemize}
-%   \item For all $\aaddr\in\dom(\msdata)$, $\pwheap(r_{\mathrm{addr}}(\aaddr)) = \stdreg{\{\aaddr\},\gc}{\lin}$ with ($\lin = \pure$ if $\aaddr\not\in A_\linear$) and ($\lin = \spatialo$ if $\aaddr\in A_\linear$)
-%   \end{itemize}
-%   with $\gc = (\ta,\stkb)$
-% \end{definition}
-% Note that for untrusted components, we require all seals to be in the corresponding code region as closure seals.
-% That is because we cannot trust untrusted components to treat their seals correctly (e.g.
-not hand them out to other components, only sign return capabilities with return seals etc.), so instead, they must be considered as closure seals, which the component is essentially allowed to sign anything with.
-
-% \begin{lemma}[World specifies linked component if it specifies each separately]
-%   \label{lem:world-specifies-linked}
-%   If
-%   \begin{itemize}
-%   \item $W \vdash_\gc \var{comp}_0$
-%   \item $W \vdash_\gc \var{comp}_0'$
-%   \end{itemize}
-%   Then $W \vdash \var{comp}_{0} \bowtie \var{comp}_{0}'$
-% \end{lemma}
-% \begin{proof}
-%   Follows easily from the definitions.
-% \end{proof}
-
-% \begin{lemma}[World specifies component pure]
-%   \label{lem:world-specifies-comp-pure}
-%   If $W \vdash_\gc \var{comp}_0$ then $\purePart{W} \vdash \var{comp}_{0}$.
-% \end{lemma}
-% \begin{proof}
-%   Follows easily from the definitions.
-% \end{proof}
-
 \subsection{Alternative formulation of reasonability?}
@@
     \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and
-      free stack $\Phi.\ms_\stk$ for all $r \not\in \{\pcreg\}$
+      free stack $\Phi.\ms_\stk$ for all $r \neq \pcreg$
     \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
@@
     behaves reasonably up to any number of steps $n$.
-    \lau{Is this reasonable as a word? In that case, it should  mention the
-      memory and free stack.}
+    % \lau{Is this reasonable as a word? In that case, it should  mention the
+    %   memory and free stack.}
+    % \dominique{No: this is reasonable as a pc.}
   \end{definition}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Clarify a proof slightly.

@@
   \label{lem:lrvtrusted-not-nequal-lrvuntrusted}
-  If $n > 0$
-  then
-  $\lrv \not\nequal \lrvg{\trusted}$
+  If $n > 0$ then $\lrv \not\nequal \lrvg{\trusted}$.
 \end{lemma}
 \begin{proof}
-  Take $\sigma \in (\gsigrets\cup \gsigcloss)$, $w = \seal{\sigma,\sigma,\sigma}$, $W$ such that $W(r) = \codereg{\sigrets,\sigcloss,\code,\gc}$, $\sigrets \subseteq \gsigrets$, $\sigcloss\subseteq \gsigcloss$, $\dom(\code) \subseteq \ta$ and $W$ empty otherwise.
+  Take $\sigma \in (\gsigrets\cup \gsigcloss)$, $w = \seal{\sigma,\sigma,\sigma}$, $W$ such that $W(r) = \codereg{\sigrets,\sigcloss,\code,\gc}$, $\sigrets \subseteq \gsigrets$, $\sigcloss\subseteq \gsigcloss$, $\sigma \in (\sigrets\cup\sigcloss)$, $\dom(\code) \subseteq \ta$ and $W$ empty otherwise.
   Then we have that $\npair[0]{(w,w)}\in \lrvg{\trusted}(W)$ but not $\npair[0]{(w,w)}\in\lrv(W)$.

### 2018-06-22 fully-abstract-comp/technical-report.tex
Simplify a proof.

@@
 
-    In this case, we have that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but we can show that it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$ by taking a $\hat{W}$ such that the following hold:
-    \begin{itemize}
-    \item $\sigrets \subseteq \gsigrets$: assumption.
-    \item $\dom(\code) \subseteq \ta$: assumption.
-    \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
-      It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions. 
-    \end{itemize}
-
-  \item $\sigma \in \sigcloss$: This follows easily from the definition of $\codereg{}$ and Lemma~\ref{lem:lrvtrusted-not-nequal-lrvuntrusted}.
+    % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$.
+    % It suffices to make choices for $b,e$ and $\hat{W}$ such that the following hold:
+    % \begin{itemize}
+    % \item $\sigrets \subseteq \gsigrets$: assumption.
+    % \item $\dom(\code) \subseteq \ta$: assumption.
+    % \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
+    %   It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions.
+    % \end{itemize}
+    %  DOMI: what if code has no call or no seal?
+    It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~ \hat{W}$.
+
+  \item $\sigma \in \sigcloss$:
+    TODO: UPDATE
+
+    This follows easily from the definition of $\codereg{}$ and Lemma~\ref{lem:lrvtrusted-not-nequal-lrvuntrusted}.
   \end{itemize}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Update one proof.

@@
     \sigrets \mathrel{\#} \sigcloss \\
-    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal}^3 \ldotp
-    \sigrets = \biguplus_{a \in \dom(\mscode)} \phi_2(d_\sigma(a)) \tand\\
-    \sigrets = \bigcup_{a \in \dom(\mscode)} \phi_1(d_\sigma(a)) \tand\\
-    \sigcloss = \bigcup_{a \in \dom(\mscode)} \phi_3(d_\sigma(a)) \tand\\
+    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
+    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
     \forall a \in \dom(\mscode) \ldotp 
-    \pi_1(d_\sigma(a)),\pi_2(d_\sigma(a)),\pi_3(d_\sigma(a)) \vdash_{\mathrm{comp-code}} \mscode,a
+    \sigrets,d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a\\
+    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
   }{
@@
 \begin{lemma}
+  \label{lem:codereg-H-not-subs-lrv}
+  If
+  \begin{itemize}
+  \item $\dom(\code) \subseteq \ta$
+  \item $\sigrets \subseteq \gsigrets$
+  \item $\sigcloss
+    \subseteq \gsigcloss$
+  \item $n > 0$
+  \item $\sigrets,\sigcloss\vdash\var{code}$
+  \end{itemize}
+  then
+  \begin{equation*}
+    \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H \not\nsubeq \stdreg{\dom(\code),\gc}{\pure}.H
+  \end{equation*}
+\end{lemma}
+\begin{proof}
+  We show that $\npair[0]{(\code,\code)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H~\hat{W}$ but not $\npair[0]{(\code,\code)} \in \stdreg{\dom(\code),\gc}{\pure}.H~\hat{W}$ for a world with only island $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}$ in $\pwfree$ but otherwise empty.
+
+  From $\sigrets,\sigcloss\vdash\var{code}$, we know that there is an $a$ such that $\code(a) = \seal{\sigma_b,\sigma_e,\_}$ with $[\sigma_b,\sigma_e] \neq \emptyset$
+  It follows easily that $\npair[0]{(\code,\code)} \not\in \stdreg{\dom(\code),\gc}{\pure}.H~\hat{W}$ since $\npair[0]{(\code(a),\code(a))} \not\in\lrv(W)$ by definition.
+
+  The fact that $\npair[0]{(\code,\code)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H~\hat{W}$ follows from Lemma~\ref{lem:ftlr-component-code-ptrs}.
+\end{proof}
+
+\begin{lemma}
   \label{lem:codereg-Hsigma-not-lrv}
-  If $\dom(\code) \subseteq \ta$,  $\sigrets \subseteq \gsigrets$, $\sigcloss \subseteq \gsigcloss$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)$.
+  If
+  \begin{itemize}
+  \item $\dom(\code) \subseteq \ta$
+  \item $\sigrets \subseteq \gsigrets$
+  \item $\sigcloss \subseteq \gsigcloss$
+  \item $n > 0$
+  \item $\sigrets,\sigcloss\vdash\var{code}$
+  \item $\sigma \in (\sigrets \uplus \sigcloss)$
+  \end{itemize}
+  then
+  \begin{equation*}
+    \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)
+  \end{equation*}
 \end{lemma}
@@
   \item $\sigma \in \sigcloss$:
-    TODO: UPDATE
-
-    This follows easily from the definition of $\codereg{}$ and Lemma~\ref{lem:lrvtrusted-not-nequal-lrvuntrusted}.
+    By definition of $\codereg{}$, it suffices to show an $\npair[0]{(\vsc,\vsc')} \in (\lrvg{\trusted}(W) \setminus \lrv(W))$ with $\exec{\vsc}$.
+    If we take $\vsc=\vsc' = ((\rx,\normal),b,e,a)$ then it suffices to show that we can take $[b,e]$ such that $[b,e] \subseteq \ta$, $\npair{[b,e]} \in\xReadCond{W}$ but not $\npair{[b,e]}\in\readCond{\normal,W}$.
+    Choosing an appropriate world, this follows by definition of $\readCond{}$ and $\xReadCond{}$ from Lemma~\ref{lem:codereg-H-not-subs-lrv}.
   \end{itemize}
@@
   \label{lem:codereg-not-read-write}
-  If $n > 0$, $\dom(\code) \subseteq \ta$, $\gc = (\ta,\stkb)$, $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and $a \in \dom(\code)$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\sigrets \neq \emptyset$, then
+  If $n > 0$, $\dom(\code) \subseteq \ta$, $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$, $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and $a \in \dom(\code)$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\sigrets \neq \emptyset$, then
   \begin{itemize}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Update and simplify a proof.

@@
     \dom(\code) = [\aaddr,\baddr] \wedge \\
-      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \trust = \trusted) \vee \\
+      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
       \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
@@
   \label{lem:codereg-not-read-write}
-  If $n > 0$, $\dom(\code) \subseteq \ta$, $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$, $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ and $a \in \dom(\code)$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\sigrets \neq \emptyset$, then
+  If
+  \begin{itemize}
+  \item $n > 0$
+  \item $\dom(\code) \subseteq \ta$
+  \item $\gc =
+    (\ta,\stkb,\gsigrets,\gsigcloss)$
+  \item $\pwheap(r_\code) \nequal
+    \codereg{\sigrets,\sigcloss,\code,\gc}$ 
+  \item $a \in \dom(\code)$
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$
+  \end{itemize}
+  then
   \begin{itemize}
@@
   \end{itemize}
-  We claim that $r_\code \in S$.
-  If this is indeed the case, then we have $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$ (we must have $R(r_\code) = \dom(\code)$ as the two regions couldn't be $\nequal$ otherwise).
-  This means that for all $a \in \dom(\code)$ and $n' < n$ it must be the case that $\npair[n']{(\code(a),\code(a))} \in \lrv(W_a)$ for all $W_a$ that are compatible w.r.t.
-  $\oplus$ and combine to be $W$.
-  We know that $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$ from which we can deduce $\npair{(\code,\code)} \in H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \xi^{-1}(W_\code)$ which in particular gives us $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code$.
-  This in turns entails that for $\sigma \in \sigrets$ there exists some $a' \in \dom(\code)$ such that $\code(a') = \seal{\sigma_\aaddr,\sigma_\baddr,\sigma_\eaddr}$ and $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
-  This means that for such a $\sigma$ and $a'$, it must be that case that
-  $\npair[n']{(\seal{\sigma_\aaddr,\sigma_\baddr,\sigma_\eaddr},\seal{\sigma_\aaddr,\sigma_\baddr,\sigma_\eaddr})}
-  \in \lrv(W_{a'})$ which gives us
-  $\codereg{\sigrets,\sigcloss,\code,\gc}.H_\sigma  \; \sigma \nequal \lrv$
-  contradicting Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
-
-  It remains to show $r_\code \in S$.
-  To show this, we assume for contradiction that $r_\code \not\in S$.
-  This means that there exists $r \neq r_\code$ such that $a \in R(r)$ and $\pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$.
   Because we know $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, it must be possible to split up $\ms_S$ in disjoint memory segments that each (paired with a target memory segment) satisfies exactly one region.
-  In this case, both $r$ and $r_\code$ needs a memory segment that has $a$ in the domain which means that no such disjoint union can exist which gives us a contradiction.
+  This means that only one region can govern address $a$, so that $r_\code \in S$
+
+  It follows that $R(r_\code) = \dom(\code)$ and $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$.
+  From the fact that $\npair{(\code,\code)} \in \pwheap(r_\code).H~W'$ for some $W'$ with $W = W' \oplus \_$, and the fact that $\dom(\code)\subseteq \ta$, it also follows that $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\sigrets,\sigcloss\vdash_{\mathrm{comp-code}} \var{code}$.
+  This now gives us a contradiction with Lemma~\ref{lem:codereg-H-not-subs-lrv}.
 \end{proof}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Refactor, simplify a proof.

@@
 \begin{lemma}
-  \label{lem:codereg-not-read-write}
-  If
-  \begin{itemize}
-  \item $n > 0$
-  \item $\dom(\code) \subseteq \ta$
-  \item $\gc =
-    (\ta,\stkb,\gsigrets,\gsigcloss)$
-  \item $\pwheap(r_\code) \nequal
-    \codereg{\sigrets,\sigcloss,\code,\gc}$ 
-  \item $a \in \dom(\code)$
-  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$
-  \end{itemize}
-  then
-  \begin{itemize}
-  \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
-    % \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
-    % It seems unnecessary to show the part about the write condition as all non-no-perm permissions have read.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  For the sake of contradiction assume $\npair{\{a\}} \in \readCond{\normal,W}$.
-  Then there exists $S \subseteq \addressable{\lin, \pwheap}$ and $R : S \fun
-  \powerset{\nats}$ such that
-  \begin{itemize}
-  \item $\biguplus_{r\in S} R(r) \supseteq A$
-  \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
-  \end{itemize}
-  Because we know $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, it must be possible to split up $\ms_S$ in disjoint memory segments that each (paired with a target memory segment) satisfies exactly one region.
-  This means that only one region can govern address $a$, so that $r_\code \in S$
-
-  It follows that $R(r_\code) = \dom(\code)$ and $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$.
-  From the fact that $\npair{(\code,\code)} \in \pwheap(r_\code).H~W'$ for some $W'$ with $W = W' \oplus \_$, and the fact that $\dom(\code)\subseteq \ta$, it also follows that $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\sigrets,\sigcloss\vdash_{\mathrm{comp-code}} \var{code}$.
-  This now gives us a contradiction with Lemma~\ref{lem:codereg-H-not-subs-lrv}.
-\end{proof}
-
-\begin{lemma} 
   \label{lem:code-reg-and-mem-sat}
@@
 $\npair[n']{R_\ms(r_{\var{code}})} \in  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc~ \xi^{-1}(R_W(r_{\var{code}}))$ as required.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:codereg-not-read-write}
+  If
+  \begin{itemize}
+  \item $n > 0$
+  \item $\dom(\code) \subseteq \ta$
+  \item $\gc =
+    (\ta,\stkb,\gsigrets,\gsigcloss)$
+  \item $\pwheap(r_\code) \nequal
+    \codereg{\sigrets,\sigcloss,\code,\gc}$
+  \item $a \in \dom(\code)$
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$
+  \end{itemize}
+  then
+  \begin{itemize}
+  \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
+    % \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
+    % It seems unnecessary to show the part about the write condition as all non-no-perm permissions have read.
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  For the sake of contradiction assume $\npair{\{a\}} \in \readCond{\normal,W}$.
+  Then there exists $S \subseteq \addressable{\lin, \pwheap}$ and $R : S \fun
+  \powerset{\nats}$ such that
+  \begin{itemize}
+  \item $\biguplus_{r\in S} R(r) \supseteq A$
+  \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
+  \end{itemize}
+  Because we know $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, it must be possible to split up $\ms_S$ in disjoint memory segments that each (paired with a target memory segment) satisfies exactly one region.
+  This means that only one region can govern address $a$, so that $r_\code \in S$
+  It follows that $R(r_\code) = \dom(\code)$ and $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$.
+
+  Lemma~\ref{lem:code-reg-and-mem-sat} tells us that $\npair{(\code,\code)} \in \pwheap(r_\code).H~W'$ for some $W'$ with $W = W' \oplus \_$.
+  Together with the fact that $\dom(\code)\subseteq \ta$, it then follows that $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\sigrets,\sigcloss\vdash_{\mathrm{comp-code}} \var{code}$.
+  This now gives us a contradiction with Lemma~\ref{lem:codereg-H-not-subs-lrv}.
 \end{proof}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Fix wrong order of arguments in seal.
Update and simplify a proof.

@@
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
     \; \middle| \;
@@
 
-\begin{lemma}
-  \label{lem:codereg-Hsigma-not-lrv}
-  If
-  \begin{itemize}
-  \item $\dom(\code) \subseteq \ta$
-  \item $\sigrets \subseteq \gsigrets$
-  \item $\sigcloss \subseteq \gsigcloss$
-  \item $n > 0$
-  \item $\sigrets,\sigcloss\vdash\var{code}$
-  \item $\sigma \in (\sigrets \uplus \sigcloss)$
-  \end{itemize}
-  then
-  \begin{equation*}
-    \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)
-  \end{equation*}
-\end{lemma}
-\begin{proof}
-  Consider the following two cases:
-  \begin{itemize}
-  \item $\sigma \in \sigrets$:
-
-    % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$.
-    % It suffices to make choices for $b,e$ and $\hat{W}$ such that the following hold:
+% \begin{lemma}
+%   \label{lem:codereg-Hsigma-not-lrv}
+%   If
 %   \begin{itemize}
-    % \item $\sigrets \subseteq \gsigrets$: assumption.
-    % \item $\dom(\code) \subseteq \ta$: assumption.
-    % \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
-    %   It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions.
+%   \item $\dom(\code) \subseteq \ta$
+%   \item $\sigrets \subseteq \gsigrets$
+%   \item $\sigcloss \subseteq \gsigcloss$
+%   \item $n > 0$
+%   \item $\sigrets,\sigcloss\vdash\var{code}$
+%   \item $\sigma \in (\sigrets \uplus \sigcloss)$
 %   \end{itemize}
-    %  DOMI: what if code has no call or no seal?
-    It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~ \hat{W}$.
-
-  \item $\sigma \in \sigcloss$:
-    By definition of $\codereg{}$, it suffices to show an $\npair[0]{(\vsc,\vsc')} \in (\lrvg{\trusted}(W) \setminus \lrv(W))$ with $\exec{\vsc}$.
-    If we take $\vsc=\vsc' = ((\rx,\normal),b,e,a)$ then it suffices to show that we can take $[b,e]$ such that $[b,e] \subseteq \ta$, $\npair{[b,e]} \in\xReadCond{W}$ but not $\npair{[b,e]}\in\readCond{\normal,W}$.
-    Choosing an appropriate world, this follows by definition of $\readCond{}$ and $\xReadCond{}$ from Lemma~\ref{lem:codereg-H-not-subs-lrv}.
-  \end{itemize}
-\end{proof}
+%   then
+%   \begin{equation*}
+%     \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)
+%   \end{equation*}
+% \end{lemma}
+% \begin{proof}
+%   Consider the following two cases:
+%   \begin{itemize}
+%   \item $\sigma \in \sigrets$:
+
+%     % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$.
+%     % It suffices to make choices for $b,e$ and $\hat{W}$ such that the following hold:
+%     % \begin{itemize}
+%     % \item $\sigrets \subseteq \gsigrets$: assumption.
+%     % \item $\dom(\code) \subseteq \ta$: assumption.
+%     % \item $\exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]}$ such that $\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen)$, $\dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr]$, $\decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$, $\code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets$.
+%     %   It is not difficult to construct a $\hat{W}$ with a suitable $\spatialo$ region to satisfy these conditions.
+%     % \end{itemize}
+%     %  DOMI: what if code has no call or no seal?
+%     It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~ \hat{W}$.
+
+%   \item $\sigma \in \sigcloss$:
+%     By definition of $\codereg{}$, it suffices to show an $\npair[0]{(\vsc,\vsc')} \in (\lrvg{\trusted}(W) \setminus \lrv(W))$ with $\exec{\vsc}$.
+%     If we take $\vsc=\vsc' = ((\rx,\normal),b,e,a)$ then it suffices to show that we can take $[b,e]$ such that $[b,e] \subseteq \ta$, $\npair{[b,e]} \in\xReadCond{W}$ but not $\npair{[b,e]}\in\readCond{\normal,W}$.
+%     Choosing an appropriate world, this follows by definition of $\readCond{}$ and $\xReadCond{}$ from Lemma~\ref{lem:codereg-H-not-subs-lrv}.
+%   \end{itemize}
+% \end{proof}
 
@@
 
-    Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\gsigrets \cup \gsigcloss)$.
-    Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.
-
-    From $\npair{(w,\_)} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.
-
-    From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.
-
-    From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
-
-    From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
-
-    From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
-    However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
+    This follows directly from $\npair{(w,\_)} \in \lrv(W_w)$ by definition of $\lrv$.
+    % Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\gsigrets \cup \gsigcloss)$.
+    % Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.
+
+    % From $\npair{(w,\_)} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.
+
+    % From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.
+
+    % From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
+
+    % From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
+
+    % From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
+    % However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
 
@@
 
+    \dominique{TODO: we don't know that a in dom(code)!}
+
     Assume that $a \in [\baddr,\eaddr] \cap \ta$ and we will prove a contradiction.
@@
 
-  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable for $n-1$ steps: By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}$, so in particular for $n-1$.
-
-    It also follows that $\sigrets \uplus \sigcloss \# \sigrets' \uplus \sigcloss'$.
+  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable for $n-1$ steps:
+
+    By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$, so in particular for $n' = n-1$.
+    It follows easily from the above that also $\npair[n-1]{(\vsc,\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W)$.
+
+    It also follows from $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, $\sigma \not\in (\gsigrets\cup \gsigcloss)$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ that $\sigrets \uplus \sigcloss \# \sigrets' \uplus \sigcloss'$.
 
@@
     The two first cases $\vsc = \retptrd(\_,\_)$ and $\vsc = \retptrc(\_,\_,\_)$ are easily discharged as the reasonability definition puts no requirements on return pointers.
-    For the final case, $\sigma \in \sigcloss$ and either
+    For the final case, $\sigma \in \sigcloss'$ and either
     \begin{itemize}
     \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n-1]{(\vsc,\_)} \in \lrv (W)$; or
-    \item $\dom(\code') \subseteq \ta \tand \sigcloss' \subseteq \gsigcloss \npair{(\vsc,\_)} \in \lrvtrusted(W)$
-    \end{itemize}
-    and in either case $\sigma \in \sigcloss'$
+    \item $\dom(\code') \subseteq \ta$, $\sigcloss' \subseteq \gsigcloss$, and $\npair{(\vsc,\_)} \in \lrvg{\trust}(W)$ for $\trust = \untrusted$ iff $\nonExec{\vsc}$.
+    \end{itemize}
     In the first case, the result follows from the induction hypothesis.
     In the second case, we have a contradiction with $\sigma \not\in (\gsigrets\cup \gsigcloss)$.
+
   \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms_S(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:

### 2018-06-22 fully-abstract-comp/technical-report.tex
Simplify lemma and proof.

@@
   \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
-  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$
-  \item $\dom(\code) \subseteq \ta$
-  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
-  \end{itemize}
-  then, with respect to $\code,\ta,\gsigrets,\gsigcloss$, $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
+  \end{itemize}
+  then, with respect to $\ta,\gsigrets,\gsigcloss$, $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
 \end{lemma}
@@
     This follows directly from $\npair{(w,\_)} \in \lrv(W_w)$ by definition of $\lrv$.
-    % Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\gsigrets \cup \gsigcloss)$.
-    % Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.
-
-    % From $\npair{(w,\_)} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.
-
-    % From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.
-
-    % From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
-
-    % From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.
-
-    % From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
-    % However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.
 
@@
 
-    By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets',\sigcloss',\mscode'$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$, so in particular for $n' = n-1$.
-    It follows easily from the above that also $\npair[n-1]{(\vsc,\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; \gc \; \sigma \; \xi^{-1}(W)$.
-
-    It also follows from $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, $\sigma \not\in (\gsigrets\cup \gsigcloss)$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ that $\sigrets \uplus \sigcloss \# \sigrets' \uplus \sigcloss'$.
+    By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets,\sigcloss,\mscode$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$, so in particular for $n' = n-1$.
+    It follows easily from the above that also $\npair[n-1]{(\vsc,\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \; \xi^{-1}(W)$.
 
@@
     The two first cases $\vsc = \retptrd(\_,\_)$ and $\vsc = \retptrc(\_,\_,\_)$ are easily discharged as the reasonability definition puts no requirements on return pointers.
-    For the final case, $\sigma \in \sigcloss'$ and either
-    \begin{itemize}
-    \item $\dom(\code') \mathrel{\#} \ta \tand \npair[n-1]{(\vsc,\_)} \in \lrv (W)$; or
-    \item $\dom(\code') \subseteq \ta$, $\sigcloss' \subseteq \gsigcloss$, and $\npair{(\vsc,\_)} \in \lrvg{\trust}(W)$ for $\trust = \untrusted$ iff $\nonExec{\vsc}$.
+    For the final case, $\sigma \in \sigcloss$ and either
+    \begin{itemize}
+    \item $\dom(\mscode) \mathrel{\#} \ta \tand \npair[n-1]{(\vsc,\_)} \in \lrv (W)$; or
+    \item $\dom(\mscode) \subseteq \ta$, $\sigcloss \subseteq \gsigcloss$, and $\npair{(\vsc,\_)} \in \lrvg{\trust}(W)$ for $\trust = \untrusted$ iff $\nonExec{\vsc}$.
     \end{itemize}
@@
       Which follows by definition of $\purePart{}$ and the fact that $R_W'(a)$ is $W_M$ with part of its ownership, but $\purePart{}$ strips away the ownership making the two compatible.
-    \item $\pwheap[W_w](r_\code) \nequal[n-1] \codereg{\sigrets,\sigcloss,\code,\gc}$
-
-      By property of $n$-equality.
-    \item $\dom(\code) \subseteq \ta$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
-
-      Follows by assumption.
-    \end{itemize}
-    which gives us that, with respect to $\code,\ta,\gsigrets,\gsigcloss$, $\pi_1(R_\ms(r))(a)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
+    \end{itemize}
+    which gives us that, with respect to $\ta,\gsigrets,\gsigcloss$, $\pi_1(R_\ms(r))(a)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
     Which is what we wanted as $\pi_1(R_\ms(r))(a) = \ms_S(a)$.
@@
       Which follows by definition of $\purePart{}$ and the fact that $R_W'(a)$ is $W_M$ with part of its ownership, but $\purePart{}$ strips away the ownership making the two compatible.
-    \item $\pwheap[W_w](r_\code) \nequal[n-1] \codereg{\sigrets,\sigcloss,\code,\gc}$
-
-      By property of $n$-equality.
-    \item $\dom(\code) \subseteq \ta$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
-
-      Follows by assumption.
-    \end{itemize}
-    which gives us that, with respect to $\code,\ta,\gsigrets,\gsigcloss$, $\pi_1(R_\ms(r))(a)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
+    \end{itemize}
+    which gives us that, with respect to $\ta,\gsigrets,\gsigcloss$, $\pi_1(R_\ms(r))(a)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$.
     Which is what we wanted as $\pi_1(R_\ms(r))(a) = \ms_\stk(a)$.

### 2018-06-22 fully-abstract-comp/technical-report.tex
Simplify consequence lemma too.
Refactor, update, simplify.

@@
     \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} c_{\mathrm{export}}}\\
-    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)
+    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
+    \dom(\msdata) \mathrel{\#} \ta
   }{
@@
     \begin{array}{l}
+      [b,e] \mathrel{\#} \ta \tand\\
       \begin{array}{r l l }
@@
     $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets
-    \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\ta)$
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\ta)$
   \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
@@
 
-\begin{lemma}
-  \label{lem:lrvtrusted-not-nequal-lrvuntrusted}
-  If $n > 0$ then $\lrv \not\nequal \lrvg{\trusted}$.
-\end{lemma}
-\begin{proof}
-  Take $\sigma \in (\gsigrets\cup \gsigcloss)$, $w = \seal{\sigma,\sigma,\sigma}$, $W$ such that $W(r) = \codereg{\sigrets,\sigcloss,\code,\gc}$, $\sigrets \subseteq \gsigrets$, $\sigcloss\subseteq \gsigcloss$, $\sigma \in (\sigrets\cup\sigcloss)$, $\dom(\code) \subseteq \ta$ and $W$ empty otherwise.
-  Then we have that $\npair[0]{(w,w)}\in \lrvg{\trusted}(W)$ but not $\npair[0]{(w,w)}\in\lrv(W)$.
-\end{proof}
-
-\begin{lemma}
-  \label{lem:codereg-H-not-subs-lrv}
-  If
-  \begin{itemize}
-  \item $\dom(\code) \subseteq \ta$
-  \item $\sigrets \subseteq \gsigrets$
-  \item $\sigcloss
-    \subseteq \gsigcloss$
-  \item $n > 0$
-  \item $\sigrets,\sigcloss\vdash\var{code}$
-  \end{itemize}
-  then
-  \begin{equation*}
-    \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H \not\nsubeq \stdreg{\dom(\code),\gc}{\pure}.H
-  \end{equation*}
-\end{lemma}
-\begin{proof}
-  We show that $\npair[0]{(\code,\code)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H~\hat{W}$ but not $\npair[0]{(\code,\code)} \in \stdreg{\dom(\code),\gc}{\pure}.H~\hat{W}$ for a world with only island $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}$ in $\pwfree$ but otherwise empty.
-
-  From $\sigrets,\sigcloss\vdash\var{code}$, we know that there is an $a$ such that $\code(a) = \seal{\sigma_b,\sigma_e,\_}$ with $[\sigma_b,\sigma_e] \neq \emptyset$
-  It follows easily that $\npair[0]{(\code,\code)} \not\in \stdreg{\dom(\code),\gc}{\pure}.H~\hat{W}$ since $\npair[0]{(\code(a),\code(a))} \not\in\lrv(W)$ by definition.
-
-  The fact that $\npair[0]{(\code,\code)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H~\hat{W}$ follows from Lemma~\ref{lem:ftlr-component-code-ptrs}.
-\end{proof}
+% \begin{lemma}
+%   \label{lem:lrvtrusted-not-nequal-lrvuntrusted}
+%   If $n > 0$ then $\lrv \not\nequal \lrvg{\trusted}$.
+% \end{lemma}
+% \begin{proof}
+%   Take $\sigma \in (\gsigrets\cup \gsigcloss)$, $w = \seal{\sigma,\sigma,\sigma}$, $W$ such that $W(r) = \codereg{\sigrets,\sigcloss,\code,\gc}$, $\sigrets \subseteq \gsigrets$, $\sigcloss\subseteq \gsigcloss$, $\sigma \in (\sigrets\cup\sigcloss)$, $\dom(\code) \subseteq \ta$ and $W$ empty otherwise.
+%   Then we have that $\npair[0]{(w,w)}\in \lrvg{\trusted}(W)$ but not $\npair[0]{(w,w)}\in\lrv(W)$.
+% \end{proof}
+
+% \begin{lemma}
+%   \label{lem:codereg-H-not-subs-lrv}
+%   If
+%   \begin{itemize}
+%   \item $\dom(\code) \subseteq \ta$
+%   \item $\sigrets \subseteq \gsigrets$
+%   \item $\sigcloss
+%     \subseteq \gsigcloss$
+%   \item $n > 0$
+%   \item $\sigrets,\sigcloss\vdash\var{code}$
+%   \end{itemize}
+%   then
+%   \begin{equation*}
+%     \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H \not\nsubeq \stdreg{\dom(\code),\gc}{\pure}.H
+%   \end{equation*}
+% \end{lemma}
+% \begin{proof}
+%   We show that $\npair[0]{(\code,\code)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H~\hat{W}$ but not $\npair[0]{(\code,\code)} \in \stdreg{\dom(\code),\gc}{\pure}.H~\hat{W}$ for a world with only island $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}$ in $\pwfree$ but otherwise empty.
+
+%   From $\sigrets,\sigcloss\vdash\var{code}$, we know that there is an $a$ such that $\code(a) = \seal{\sigma_b,\sigma_e,\_}$ with $[\sigma_b,\sigma_e] \neq \emptyset$
+%   It follows easily that $\npair[0]{(\code,\code)} \not\in \stdreg{\dom(\code),\gc}{\pure}.H~\hat{W}$ since $\npair[0]{(\code(a),\code(a))} \not\in\lrv(W)$ by definition.
+
+%   The fact that $\npair[0]{(\code,\code)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H~\hat{W}$ follows from Lemma~\ref{lem:ftlr-component-code-ptrs}.
+% \end{proof}
 
@@
 
-\begin{lemma}
-  \label{lem:codereg-not-read-write}
-  If
-  \begin{itemize}
-  \item $n > 0$
-  \item $\dom(\code) \subseteq \ta$
-  \item $\gc =
-    (\ta,\stkb,\gsigrets,\gsigcloss)$
-  \item $\pwheap(r_\code) \nequal
-    \codereg{\sigrets,\sigcloss,\code,\gc}$
-  \item $a \in \dom(\code)$
-  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$
-  \end{itemize}
-  then
-  \begin{itemize}
-  \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
-    % \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
-    % It seems unnecessary to show the part about the write condition as all non-no-perm permissions have read.
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  For the sake of contradiction assume $\npair{\{a\}} \in \readCond{\normal,W}$.
-  Then there exists $S \subseteq \addressable{\lin, \pwheap}$ and $R : S \fun
-  \powerset{\nats}$ such that
-  \begin{itemize}
-  \item $\biguplus_{r\in S} R(r) \supseteq A$
-  \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
-  \end{itemize}
-  Because we know $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, it must be possible to split up $\ms_S$ in disjoint memory segments that each (paired with a target memory segment) satisfies exactly one region.
-  This means that only one region can govern address $a$, so that $r_\code \in S$
-  It follows that $R(r_\code) = \dom(\code)$ and $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$.
-
-  Lemma~\ref{lem:code-reg-and-mem-sat} tells us that $\npair{(\code,\code)} \in \pwheap(r_\code).H~W'$ for some $W'$ with $W = W' \oplus \_$.
-  Together with the fact that $\dom(\code)\subseteq \ta$, it then follows that $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\sigrets,\sigcloss\vdash_{\mathrm{comp-code}} \var{code}$.
-  This now gives us a contradiction with Lemma~\ref{lem:codereg-H-not-subs-lrv}.
-\end{proof}
+% \begin{lemma}
+%   \label{lem:codereg-not-read-write}
+%   If
+%   \begin{itemize}
+%   \item $n > 0$
+%   \item $\dom(\code) \subseteq \ta$
+%   \item $\gc =
+%     (\ta,\stkb,\gsigrets,\gsigcloss)$
+%   \item $\pwheap(r_\code) \nequal
+%     \codereg{\sigrets,\sigcloss,\code,\gc}$
+%   \item $a \in \dom(\code)$
+%   \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$
+%   \end{itemize}
+%   then
+%   \begin{itemize}
+%   \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
+%     % \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
+%     % It seems unnecessary to show the part about the write condition as all non-no-perm permissions have read.
+%   \end{itemize}
+% \end{lemma}
+% \begin{proof}
+%   For the sake of contradiction assume $\npair{\{a\}} \in \readCond{\normal,W}$.
+%   Then there exists $S \subseteq \addressable{\lin, \pwheap}$ and $R : S \fun
+%   \powerset{\nats}$ such that
+%   \begin{itemize}
+%   \item $\biguplus_{r\in S} R(r) \supseteq A$
+%   \item $\forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$
+%   \end{itemize}
+%   Because we know $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, it must be possible to split up $\ms_S$ in disjoint memory segments that each (paired with a target memory segment) satisfies exactly one region.
+%   This means that only one region can govern address $a$, so that $r_\code \in S$
+%   It follows that $R(r_\code) = \dom(\code)$ and $\pwheap(r_\code).H \nsubeq \stdreg{\dom(\code),\gc}{\pur}.H$.
+
+%   Lemma~\ref{lem:code-reg-and-mem-sat} tells us that $\npair{(\code,\code)} \in \pwheap(r_\code).H~W'$ for some $W'$ with $W = W' \oplus \_$.
+%   Together with the fact that $\dom(\code)\subseteq \ta$, it then follows that $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$ and $\sigrets,\sigcloss\vdash_{\mathrm{comp-code}} \var{code}$.
+%   This now gives us a contradiction with Lemma~\ref{lem:codereg-H-not-subs-lrv}.
+% \end{proof}
 
@@
 
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$, then $[\baddr,\eaddr] \# \ta$:
-
-    \dominique{TODO: we don't know that a in dom(code)!}
-
-    Assume that $a \in [\baddr,\eaddr] \cap \ta$ and we will prove a contradiction.
-
-    From $\npair{(w,\_)} \in \lrv(W_w)$, it follows that $\npair{\{a\}}\in\readCond{\normal,W_w}$.
-
-    This contradicts Lemma~\ref{lem:codereg-not-read-write}.
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \# \ta$:
+
+    This follows directly from $\npair{(w,\_)} \in \lrv(W_w)$ by definition of $\lrv$.
 
@@
   \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined
-  \item $\pwheap[W_w](r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
-  \item $\dom(\code) \subseteq \ta$
-  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
-  \end{itemize}
-  then, with respect to $\code,\ta,\gsigrets,\gsigcloss$,
+  \end{itemize}
+  then, with respect to $\ta,\gsigrets,\gsigcloss$,
   \begin{itemize}
@@
 \begin{proof}
-  We prove the three results separately:
+  We prove the two results separately:
   \begin{itemize}
@@
   \item $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$, with respect to $\code,\ta,\gsigrets,\gsigcloss$.
-  \item $\gsigrets \neq \emptyset$
   \item $n > 0$
@@
     The former follows directly from Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
-    The latter follows by Theorem~\ref{thm:ftlr}, Lemma~\ref{lem:conds-shrinkable} and definition of $\execCond{}$.
+    The latter follows by Theorem~\ref{thm:ftlr}, Lemmas~\ref{lem:conds-shrinkable} and~\ref{lem:monotonicity} and definition of $\execCond{}$.
   \end{itemize}
@@
   \begin{itemize}
-  \item $\trust = \untrusted$ and $\dom(\mscode \mathrel{\#} \ta)$:
+  \item $\trust = \untrusted$ and $\dom(\mscode) \mathrel{\#} \ta$:
     By definition of $\lrv(W)$, it suffices to show that:
     \begin{itemize}
+    \item $[b,e] \mathrel{\#} \ta$: follows directly from $[\baddr,\eaddr]\subseteq \dom(\mscode)$ and $\dom(\mscode) \mathrel{\#} \ta$.
     \item $\npair{[b,e]}\in\readCond{\normal,W}$:
@@
       \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
-    \end{equation*}
-    Using the former, the result follows easily by definition.
+    \end{equation*}:
+
+    Using $\mscode(a) \in \ints$, the result follows easily by definition.
   \end{itemize}
@@
   \item $w \not\in \overline{c_{\mathrm{import}}}$
+  \item $(A_{\mathrm{own}} \cup A_{\mathrm{non-linear}}) \mathrel{\#} \ta$
   \end{itemize}
@@
 
-    By definition of $\lrv$, it suffices to prove that $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ and $\npair{[\baddr,\eaddr]}\in\writeCond{\lin,W}$.
+    By definition of $\lrv$, it suffices to prove that $[b,e] \mathrel{\#} \ta$, $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W}$ and $\npair{[\baddr,\eaddr]}\in\writeCond{\lin,W}$.
     The result then follows easily from the assumptions, with Lemma~\ref{lem:conds-shrinkable}.
@@
   \item If $w = \sealed{\sigma,\vsc}$ with $\sigma \in \sigcloss$ and $\exec{\vsc}$, then
+  \item $A_{\mathrm{non-linear}}\mathrel{\#} \ta$
     $\vsc$ behaves reasonably up to $n$ steps.
@@
 
-  \item TODO: the other case...
+  \item
+    \begin{equation*}
+      \dom(\mscode),\emptyset,A_{\mathrm{non-linear}},\emptyset,\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+    \end{equation*}
+
+    In this case, the result follows from Lemma~\ref{lem:ftlr-comp-value}, using the fact that $\readCond$ and $\writeCond$ follow trivially for the empty $A_{\mathrm{own}}$ and a similar observation about the empty imports.
   \end{itemize}
@@
   \item $(\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)$
+  \item $\dom(\msdata) \mathrel{\#} \ta$
   \end{itemize}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Remove obsolete comment.
Discharge a technical proof.
Standardize naming a bit.

@@
 \label{sec:full-abstraction}
-\lau{TODOs from last day in Leuven}
 \lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}
@@
   \label{lem:unique-h-sigma}
-  if $\iota,iota' \in \dom(\pwheap)$ and $\pwheap(\iota) = (\pure,\_,H_\sigma)$
-  and $\pwheap(\iota') = (\pure,\_,H_\sigma')$ and $\sigma \in \dom(H_\sigma)$
+  If $r,r' \in \dom(\pwheap)$ and $\pwheap(r) = (\pure,\_,H_\sigma)$
+  and $\pwheap(r') = (\pure,\_,H_\sigma')$ and $\sigma \in \dom(H_\sigma)$
   and $\sigma \in \dom(H_\sigma')$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$,
@@
   \[
-    H_\sigma = H_\sigma' \text{ and } \iota = \iota'
+    H_\sigma = H_\sigma' \text{ and } r = r'
   \]
@@
 \begin{proof}
-  \lau{TODO}
+  This follows easily by definition of $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$ and $\lrheap$.
 \end{proof}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Add comment.

@@
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to
-      $n'-i - 1$ steps
+    \item $\Phi'(\pcreg) + \calllen$ behaves reasonably up to $n'-i - 1$ steps
     \end{itemize}
@@
 
+\dominique{Why is the below lemma needed?}
 \begin{lemma}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Remove unnecessary requirement.
Some updates to full abstraction.

--- 2018-06-22 fully-abstract-comp/technical-report.tex
@@
     $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\ta)$
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \# \dom(\ta)$
   \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
@@
     behaves reasonably up to any number of steps $n$.
+
+    We say that a component
+    $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is reasonable if $\var{comp}_0$ is reasonable.
     % \lau{Is this reasonable as a word? In that case, it should  mention the
@@
 \begin{lemma}[FTLR for components]
+  \label{lem:ftlr-comps}
   If
   \begin{itemize}
-  \item $\gc = (\ta,\stkb)$
+  \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
   \item $\var{comp}$ is a well-formed component, i.e. $\vdash \var{comp}$
@@
     \item $\var{comp}$ is a reasonable component (see
-      Section~\ref{sec:reas-comp}) and $\dom(\var{comp}.\mscode) \subseteq \ta$
+      Section~\ref{sec:reasonability}) and $\dom(\var{comp}.\mscode) \subseteq \ta$
     \item $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
@@
   \begin{equation*}
-    \forall \src{\context} \ldotp \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
-  \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$
+    \forall \src{\context} \ldotp \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1},\gsigrets_1,\gsigcloss_1} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2,\gsigrets_2,\gsigcloss_2}}
+  \end{equation*}
+  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and $\gsigcloss_i = \var{comp}_i.\sigcloss$
+  \dominique{and stkbase?}
 \end{definition}
@@
   \begin{equation*}
-    \forall \trg{\context} \ldotp \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
+    \forall \trg{\context} \ldotp \vdash \trg{\context} \Rightarrow \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
   \end{equation*}
@@
   \label{thm:full-abstraction}
-  For reasonable components $\var{comp}_1$ and $\var{comp}_2$, we have
+  For reasonable, well-formed components $\var{comp}_1$ and $\var{comp}_2$, we have
   \begin{gather*}

### 2018-06-22 fully-abstract-comp/ftlr-proof-cases.tex
Remove obsolete TODOs.
compilation issue
Fix a cross-cutting problem related to a bad definition of linking.
The problem was that linking of components was poorly designed, with components
already expected to already include the exports of the components they link
with, rather than them being inserted by the linking process.
This worked in principle, except that a single component could not be linked
with two different other components (as in the statement of full abstraction).
The fix is to introduce a more reasonable linking setup: components specify
imports as symbol-address pairs indicatin

--- 2018-06-21 fully-abstract-comp/ftlr-proof-cases.tex
@@
     \begin{itemize}
-    \item $\Phi_S(\rstk) \neq \stkptr{\_,\_,\_,\_}$\\
+    \item $\Phi_S(\rstk) \neq \stkptr{\_,\_,\_,\_}$
+
       In this case due to $\reg[3]_S$ being related to $\reg[3]_T$, there are three cases we need to consider.
@@
   \item $\purePart{W_w} \oplus \purePart{W_M}$ is defined: By Lemma~\ref{lem:purePart-oplus}
-  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,\gc}$: TODO
-  \item $\dom(\code) \subseteq \ta$: Follows from $\trust = \trusted$ by the choice of $\trust$.
-  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$: TODO
   \end{itemize}

### 2018-06-22 fully-abstract-comp/technical-report.tex
Fix a cross-cutting problem related to a bad definition of linking.
The problem was that linking of components was poorly designed, with components
already expected to already include the exports of the components they link
with, rather than them being inserted by the linking process.
This worked in principle, except that a single component could not be linked
with two different other components (as in the statement of full abstraction).
The fix is to introduce a more reasonable linking setup: components specify
imports as symbol-address pairs indicating the address for the symbol should be
su

@@
     \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
-    &\Word & \defeq & \ints \uplus \Caps\\
+    w \in &\Word & \defeq & \ints \uplus \Caps\\
     \permbnf \in& \Perm & \defbnf & \dots \\
@@
 \begin{align*}
+  s &\in \Symbol\\
+  \var{import} &\mathrel{::=} a \mapsfrom s\\
+  \var{export} &\mathrel{::=} s \mapsto w\\
   \var{comp}_0 &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)\\
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
-  }
-  \and
-  \inference{
-    c \in \overline{c_{\mathrm{import}}}
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} z
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
   }
@@
   \inference{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc \\
     \sigma \in \sigcloss
   }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \sealed{\sigma,\vsc}
   }
@@
   }{
-    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-export}} \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
   }
@@
   \inference{
-    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\emptyset,\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
   }{
-    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
   }
@@
     A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
-    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} c_{\mathrm{export}}}\\
-    \overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
+    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
+    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} & \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}}& \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
+    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}\\
+    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
     \dom(\msdata) \mathrel{\#} \ta
   }{
-    \vdash (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
+    \vdash (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
   }
@@
   \inference{
-    \var{comp}_0 = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)\\
-    \vdash \var{comp}_0 & c_{\mathrm{main},c}, c_{\mathrm{main},d} \in \overline{c_{\mathrm{export}}}
+    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
+    \vdash \var{comp}_0 & (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
   }{
@@
     \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
-    \mscode[3] = \mscode[1] \uplus \mscode[2] &
-    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
-    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}&
-    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}\\
+    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
+    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
+    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
+    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
     \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
@@
       [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
-      \overline{c_{\mathrm{import}}} = \emptyset
+      \overline{\var{import}} = \emptyset
     }{
-      ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
+      ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
     }
@@
     We say that a component
-    $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$
-    is reasonable if the following hold: For all $\sealed{\sigma,\vsc} \in
+    $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$
+    is reasonable if the following hold: For all $(s \mapsto \sealed{\sigma,\vsc}) \in
     \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$, we have that $\vsc$
@@
      &\npair{\var{comp},\var{comp}} \;\mid \;\\
-      &\qquad\var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+      &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
       &\qquad\text{For all } W' \future W \ldotp \\
-      &\qquad\quad\text{If } \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
-      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W') \tand\\
-      &\qquad\quad\overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+      &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+      &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
+      &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
   \end{aligned}
@@
      \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
-     \{c_{\mathrm{main},c},c_{\mathrm{main},d}\} \subseteq \overline{c_{\mathrm{export}}}
+     \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
        \end{multlined}
@@
   \begin{itemize}
-  \item $\var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1],A_{\linear,1})$
-  \item $\var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2],A_{\linear,2})$
-  \item $\var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3],A_{\linear,3})$
+  \item $\var{comp}_1 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})$
+  \item $\var{comp}_2 = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})$
+  \item $\var{comp}_3 = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})$
   \item $\mscode[3] = \mscode[1] \uplus \mscode[2]$
-  \item $\msdata[3] = \msdata[1] \uplus \msdata[2]$
-  \item $\overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}$
-  \item $\overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}$
+  \item $\msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3]$
+  \item $\overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}$
+  \item $\overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}$
   \item $\sigrets[3] = \sigrets[1] \uplus \sigrets[2]$
@@
 
-  Now take $W' \future (W_1\uplus W_2)$ and assume that $\overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
+  Now take $W' \future (W_1\uplus W_2)$ and assume that $\overline{\npair[n']{(w_{\mathrm{import},3},w_{\mathrm{import},3})}} \in \lrv(\purePart{W'})$ for all $n' < n$.
+  Take $\overline{\var{import}_3} = \overline{a_{\mathrm{import},3} \mapsfrom s_{\mathrm{import},3}}$.
+  Take $\msdata[3]' = \msdata[3]{}[\overline{a_{\mathrm{import},3} \mapsto w_{\mathrm{import}}}]$
   Then it remains to show that
   \begin{equation*}
-    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3], \mscode[3]\uplus\msdata[3])} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
   \end{equation*}
@@
 
-  First, we prove that
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import},1},c_{\mathrm{import},1})}} \in \lrv(\purePart{W'})
-  \end{equation*}
-  and
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import},2},c_{\mathrm{import},2})}} \in \lrv(\purePart{W'})
+  First, note that
+  \begin{align*}
+    \msdata[3]' &= \msdata[3]{}[\overline{a_{\mathrm{import},3} \mapsto w_{\mathrm{import}}}]\\
+                &=(\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3]\\
+                  &\hspace{4cm}[\overline{a_{\mathrm{import},3} \mapsto w_{\mathrm{import}}}]\\
+                &= \msdata[1][a \mapsto w \mid (a \mapsfrom s) \in \overline{\var{import}_1}, (s \mapsto w) \in \overline{\var{export}}_2][\overline{a_{\mathrm{import},3} \mapsto w_{\mathrm{import},3}} \mid \overline{a_{\mathrm{import},3}\mapsfrom \_} \in \overline{\var{import}_1}]\\
+                &\uplus \msdata[2][a \mapsto w \mid (a \mapsfrom s) \in \overline{\var{import}_2}, (s \mapsto w) \in \overline{\var{export}}_1][\overline{a_{\mathrm{import},3} \mapsto w_{\mathrm{import},3}} \mid \overline{a_{\mathrm{import},3}\mapsfrom \_} \in \overline{\var{import}_2}]\\
+                &= \msdata[1]' \uplus \msdata[2]'
+  \end{align*}
+
+  First, we prove that for all substituted values $w$ in the equations above, we have that
+  \begin{equation*}
+    \overline{\npair[n']{(w,w)}} \in \lrv(\purePart{W'})
   \end{equation*}
   for all $n' < n$.
-  Simple set algebra tells us that 
-  \begin{equation*}
-    \overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}} \subseteq \overline{c_{\mathrm{export},3}} \cup \overline{c_{\mathrm{import},3}}
-  \end{equation*}
-  and for all $n' < n$, we know that
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import},3},c_{\mathrm{import},3})}} \in \lrv(\purePart{W'})
-  \end{equation*}
-  and
-  \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{export},3},c_{\mathrm{export},3})}} \in \lrv(\purePart{W'})\text{.}
-  \end{equation*}
-
-  Then, it follows from our assumptions ($\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$) that
+  We know by assumption that this is true for the $\overline{w_{\mathrm{import}}}$ such that $\overline{a_{\mathrm{import},3}\mapsfrom \_} \in \overline{\var{import}_1}$ or $\overline{a_{\mathrm{import},3}\mapsfrom \_} \in \overline{\var{import}_2}$.
+  On the other hand, we know by induction that this is true for the $w$ such that $(s \mapsto w) \in \overline{\var{export}}_1$ or $(s \mapsto w) \in \overline{\var{export}}_2$.
+
+  Then, it follows from our assumptions ($\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp(W_1)$ and $\npair{(\var{comp}_2,\var{comp}_2)} \in \lrcomp(W_2)$), and using Lemma~\ref{lem:uplus-future} that
   \begin{align*}
-    \npair{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1], \mscode[1]\uplus\msdata[1])} &\in \lrheap(\pwheap[W_1])(W')\\
-    \npair{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2], \mscode[2]\uplus\msdata[2])} &\in \lrheap(\pwheap[W_2])(W')
+    \npair{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1]', \mscode[1]\uplus\msdata[1]')} &\in \lrheap(\pwheap[W_1])(W')\\
+    \npair{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2]', \mscode[2]\uplus\msdata[2]')} &\in \lrheap(\pwheap[W_2])(W')
   \end{align*}
@@
   \begin{equation*}
-    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3], \mscode[3]\uplus\msdata[3])} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
   \end{equation*}
@@
   \begin{equation*}
-    \dom(\mscode), A_{\mathrm{own}}, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+    \dom(\mscode), A_{\mathrm{own}}, A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
   \end{equation*}
@@
   \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$
-  \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(\purePart{W})$ for all $n' < n$
-  \item $w \not\in \overline{c_{\mathrm{import}}}$
   \item $(A_{\mathrm{own}} \cup A_{\mathrm{non-linear}}) \mathrel{\#} \ta$
@@
   \item $w = z$: result follows trivially
-  \item $w = c$, $c \in \overline{c_{\mathrm{import}}}$: contradicts assumption.
   \item $w = ((\perm,\lin),\baddr,\eaddr,\aaddr)$,
@@
     \begin{equation*}
-      \dom(\mscode),A_{\mathrm{own}},A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc
+      \dom(\mscode),A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \vsc
     \end{equation*}
@@
 
-    First, we have that $\npair{(\vsc,\vsc)} \in \lrv (W)$.
-    This follows by induction (if $\vsc \not\in \overline{c_{\mathrm{import}}}$) or by assumption (if $\vsc \in \overline{c_{\mathrm{import}}}$).
+    First, we have by induction that $\npair{(\vsc,\vsc)} \in \lrv (W)$.
 
@@
   \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$
-  \item If $w = \sealed{\sigma,\vsc}$ with $\sigma \in \sigcloss$ and $\exec{\vsc}$, then
+  \item If $w = \sealed{\sigma,\vsc}$ with $\dom(\mscode) \subseteq \ta$, $\sigma \in \sigcloss$ and $\exec{\vsc}$, then
+    $\vsc$ behaves reasonably up to $n$ steps.
   \item $A_{\mathrm{non-linear}}\mathrel{\#} \ta$
-    $\vsc$ behaves reasonably up to $n$ steps.
   \end{itemize}
@@
 
-      From one of our assumptions, we know that $\vsc$ behaves reasonably up to $n$ steps.
+      From one of our assumptions, we know that $\vsc$ behaves reasonably up to $n$ steps if $\dom(\mscode)\subseteq \ta$.
 
@@
     \begin{equation*}
-      \dom(\mscode),\emptyset,A_{\mathrm{non-linear}},\emptyset,\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
+      \dom(\mscode),\emptyset,A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} w
     \end{equation*}
@@
     \begin{itemize}
-    \item $\var{comp}$ is a reasonable component (see
-      Section~\ref{sec:reasonability}) and $\dom(\var{comp}.\mscode) \subseteq \ta$
+    \item $\dom(\var{comp}.\mscode) \subseteq \ta$ and $\var{comp}$ is a reasonable component (see Section~\ref{sec:reasonability})
     \item $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
@@
   \begin{equation*}
-    \var{comp} = (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)
+    \var{comp} = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
   \end{equation*}
@@
     \begin{equation*}
-      \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)
+      \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)
     \end{equation*}
@@
     \begin{equation*}
-      \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}
+      \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} c_{\mathrm{export}}}
     \end{equation*}
-  \item $\overline{c_{\mathrm{export}}} \mathrel{\#} \overline{c_{\mathrm{import}}}$
   \item $(\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)$
@@
 
-  Take $W$ such that
+  Take $\overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}}$.
+
+  Now take $W$ such that
   \begin{itemize}
@@
   \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$} 
+    \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+  \end{equation*}
+  and
+  \begin{equation*}
+    \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
   \end{equation*}
@@
     \begin{equation*}
-      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
+      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
     \end{equation*}
@@
     \begin{itemize}
-    \item $\npair{(\msdata|_{\{\addr\}},\msdata|_{\{\addr\}})} \in \stdreg{\{\aaddr\},\gc}{\lin}.H~ \xi^{-1}(R_W(r_\addr))$:
+    \item $\npair{(\msdata'|_{\{\addr\}},\msdata'|_{\{\addr\}})} \in \stdreg{\{\aaddr\},\gc}{\lin}.H~ \xi^{-1}(R_W(r_\addr))$:
       By definition, it suffices to prove that for all $n' < n$,
-      $\npair[n']{(\msdata(\addr),\msdata(\addr))} \in \lrv(R_W(r_\addr))$.
-
-      We can assume that $\msdata(\addr) \not\in \overline{c_{\var{import}}}$, because otherwise the result follows from the previous assumption that 
-      \begin{equation*}
-        \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$} 
+      $\npair[n']{(\msdata'(\addr),\msdata'(\addr))} \in \lrv(R_W(r_\addr))$.
+
+      If $(a \mapsfrom \_) \in \overline{\var{import}}$, then we know that $\msdata'(a)$ is the corresponding $w_{\mathrm{import}}$ and the result is fine by the assumption that
+      \begin{equation*}
+        \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
       \end{equation*}
@@
 
-      We know from our assumptions that
-      \begin{equation*}
-        \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\overline{c_{\mathrm{import}}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)
+      Otherwise, $\msdata'(a) = \msdata(a)$ and we know from our assumptions that
+      \begin{equation*}
+        \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)
       \end{equation*}
@@
       \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$: by assumption.
-      \item $\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})} \in \lrv(\purePart{R_W(\addr)})$ for all $n' < n$: by the fact that $\purePart{R_W(\addr)} = \purePart{W'}$ and the above assumption that .
-        \begin{equation*}
-          \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$} 
-        \end{equation*}
-      \item $w \not\in \overline{c_{\mathrm{import}}}$: by assumption.
       \end{itemize}
@@
   \end{itemize}
-  \item $\overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})$:
+  \item $\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})$:
 
@@
     \begin{equation*}
-      \overline{\dom(\mscode), \emptyset, A_{\mathrm{non-linear}}, \overline{c_{\mathrm{import}}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-value}} c_{\mathrm{export}}}
+      \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
     \end{equation*}
@@
     \item $\dom(\mscode) \mathrel{\#} \ta$ or $\dom(\mscode) \subseteq \ta$: by assumption.
+    \item If $w_{\mathrm{export}} = \sealed{\sigma,\vsc}$ with $\dom(\mscode) \subseteq \ta$, $\sigma \in \sigcloss$ and $\exec{\vsc}$, then $\vsc$ behaves reasonably up to $n$ steps:
+
+      This follows directly from the fact that $\var{comp}$ is a reasonable component.
+
+    \item $A_{\mathrm{non-linear}}\mathrel{\#} \ta$:
+      This follows from the facts that $A_{\mathrm{non-linear}}\subseteq \dom(\msdata) \mathrel{\#} \ta$.
     \end{itemize}
@@
 \begin{proof}
-  Take $\var{comp} = ((\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
+  Take $\var{comp} = ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d})$.
   Then we know from $\var{comp} \rightsquigarrow \Phi$ that
@@
   \item $[\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata))$
-  \item $\overline{c_{\mathrm{import}}} = \emptyset$
+  \item $\overline{\var{import}} = \emptyset$
   \item $\Phi = (\mem, \reg\src{, \emptyset, \ms_\stk})$
@@
   From $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$, we know that
-  $\{c_{\mathrm{main},c}, c_{\mathrm{main},d}\}\subseteq \overline{c_{\mathrm{export}}}$ and 
-  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{c_{\mathrm{import}}}, \overline{c_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
+  $\{(\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d})\}\subseteq \overline{\var{export}}$ and 
+  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{a_{\mathrm{import}}\mapsfrom s_{\mathrm{import}}}, \overline{s_{\mathrm{export}}\mapsto w_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
   From this, it follows that for all  $W' \future W$, if
   \begin{equation*}
-    \overline{\npair[n']{(c_{\mathrm{import}},c_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
+    \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}
   \end{equation*}
+  and $\msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}]$
   then
   \begin{equation*}
-    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
+    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
   \end{equation*}
@@
   \begin{equation*}
-    \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+    \overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
   \end{equation*}
-  Since $\overline{c_{\mathrm{import}}} = \emptyset$, the former holds vacuously and the latter two results follow for every $W' \future W$.
+  Since $\overline{w_{\mathrm{import}}} = \emptyset$, the former holds vacuously, $\msdata' = \msdata$ and the latter two results follow for every $W' \future W$.
 
@@
     \begin{equation*}
-      \overline{\npair{(c_{\mathrm{export}},c_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+      \overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
     \end{equation*}
@@
 
-    We also know that $c_{\mathrm{main},c} = \sealed{\sigma, c_{\mathrm{main},c}'}$, $c_{\mathrm{main},d} = \sealed{\sigma, c_{\mathrm{main},d}'}$, $\sigma=\sigma_1 = \sigma_2$, so by definition of $\lrv$, we know that there exist $r_1,r_2 \in \dom(\pwheap[W_\pc])$ such that $\pwheap[W_\pc](r_1) = (\pure,\_,H_{\sigma,1})$, $\pwheap[W_\pc](r_2) = (\pure,\_,H_{\sigma,2})$ and
-    \begin{align*}
-      \npair[n']{(c_{\mathrm{main},d}',c_{\mathrm{main},d}')} &\in H_{\sigma,1}~ \sigma~ \xi^{-1}(W_\pc) \text{ for all } n' < n\\
-      \npair[n']{(c_{\mathrm{main},d}',c_{\mathrm{main},d}')} &\in H_{\sigma,2}~ \sigma~ \xi^{-1}(W_\pc) \text{ for all } n' < n
-    \end{align*}
-    By definition of $W_\stk$, we know that it contains no regions that define $H_\sigma$.
-    Therefore, using the definition of $W'$, we can deduce that $r_1,r_2 \in \dom(\pwheap)$.
-    From the fact that 
-    \begin{equation*}
-      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap)(W')
+    Since $W_\pc = \purePart{W'}$, Lemma~\ref{lem:safe-vals-safe-exec} tells us that for $n' < n$,
+    \begin{equation*}
+      \npair[n']{( (c_{\mathrm{main},c},c_{\mathrm{main},d}), (c_{\mathrm{main},c},c_{\mathrm{main},d}) )} \in \lrexj(W_\pcreg)
     \end{equation*}
-    it then follows that $\sigma$ can only be defined in $H_\sigma$ for only one region, so that $r_1 = r_2 = r$ and $H_{\sigma,1} = H_{\sigma,2} = H_\sigma$.
-
-    Since $W_\pc = \purePart{W'}$, we have (in both the linear and non-linear case), that for $n' < n$, $\npair[n']{( (c_{\mathrm{main},c},c_{\mathrm{main},d}), (c_{\mathrm{main},c},c_{\mathrm{main},d}) )} \in \lrexj(W_\pcreg)$.
     Since $\reg(\pcreg) = c_{\mathrm{main},c}'$ and $\reg(\rdata) = c_{\mathrm{main},d}'$, this is what we set out to prove.

### 2018-06-22 fully-abstract-comp/technical-report.tex
Prove full abstraction.

@@
 \begin{lemma}[Adequacy of execution configuration LR]
+  \label{lem:adequacy}
   If
   \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq](W)$
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$
   \item $i \leq n$
@@
   \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq](W)$
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$
   \item $i \leq n$
@@
 \begin{lemma}[Compatibility lemma for context plugging]
+  \label{lem:compat-context-plug}
   If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
@@
 \begin{proof}
-  First, 
-  TODO
+  \begin{itemize}
+  \item Consider first the upward arrow.
+    Assume $\comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}$.
+
+    Take a $\src{\context}$ such that $\vdash \src{\context}$, take $\src{\ta[,i]}
+    = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and
+    $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$ and we will prove that
+    $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}} \Leftrightarrow
+    \src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
+
+    By symmetry, we can assume w.l.o.g. that $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$ and prove that $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
+    Note that this implies that $\src{\context}$ is a valid context for both $\src{\var{comp}_1}$ and $\src{\var{comp}_2}$.
+
+    First, we show that also $\plug{\context}{\var{comp}_1} \trg{\term}$.
+    Take $n$ the amount of steps in the termination of $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\preceq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\preceq,\gc_1](W_1\uplus W_1')$
+    Together with $\src{\plug{\context}{\var{comp}_1} \sterm[n]{\gc_1}}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \trg{\term}$.
+
+    It follows from $\comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}$ that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
+
+    It now remains to show that also $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
+    Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \trg{\term}$.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\succeq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\succeq,\gc_2](W_2\uplus W_2')$
+    Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$, concluding this direction of the proof.
+
+  \item The downward arrow is similar.
+
+    Assume $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$.
+    Take $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$.
+
+    Take a $\trg{\context}$ such that $\vdash \trg{\context}$ and we will prove that
+    $\trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow
+    \trg{\plug{\context}{\var{comp}_2} \term}$.
+
+    By symmetry, we can assume w.l.o.g. that $\trg{\plug{\context}{\var{comp}_1} \term}$ and prove that $\trg{\plug{\context}{\var{comp}_2} \term}$.
+    Note that this implies that $\trg{\context}$ is a valid context for both $\trg{\var{comp}_1}$ and $\trg{\var{comp}_2}$.
+
+    First, we show that also $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
+    Take $n$ the amount of steps in the termination of $\plug{\context}{\var{comp}_1} \trg{\term}$.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
+    Together with $\plug{\context}{\var{comp}_1} \trg{\term[n]}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
+
+    It follows from $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
+
+    It now remains to show that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
+    Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$
+    Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_2} \trg{\term}$, concluding the second direction of the proof.
+  \end{itemize}
 \end{proof}

### 2018-06-24 fully-abstract-comp/ftlr-proof-cases.tex
Typos, errors from reformatting FTLR proof, and formatting.

--- 2018-06-22 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \item $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$
-  \item $\memSat{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$
+  \item $\memSat{\Phi_S.\mem,\Phi_S.\stk,\Phi_S.\ms_\stk,\Phi_T.\mem}{W_M}$
   \item $W_\pcreg \oplus W_R \oplus W_M$ is defined.
@@
   \[
-    \npair[n'-1]{(\Phi_S',\Phi_T')} \in \lro
+    \npair[n-1]{(\Phi_S',\Phi_T')} \in \lro
   \]
@@
     \begin{itemize}
-    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n'-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair[n'-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
     \end{itemize}: Follows from the corresponding assumption of this lemma and
@@
     Follows from the definition of $\updPcAddr{}$ and the assumptions of this case.
-  \item $\npair[n'-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
     Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}.
-  \item $\npair[n' - 1]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
+  \item $\npair[n-1]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
     Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed} and the fact that integers are always in $\lrvg{\trust}$ by definition.
-  \item $\memSat[n'-1]{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
+  \item $\memSat[n-1]{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
     Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}.
-  \item Theorem~\ref{thm:ftlr} holds for all $n'' < n'-1$:
-    follows from the corresponding assumption of this lemma, since $n' - 1 < n$.
-  \end{itemize}
-
-  \lau{TODO: look through this case and find the place where the ''no seal cross
-    boundary thing'' should have been used. (I have done this, and it does not
-    seem necessary (?))}
+  \item Theorem~\ref{thm:ftlr} holds for all $n'' < n-1$:
+    follows from the corresponding assumption of this lemma, since $n - 1 < n$.
+  \end{itemize}
+
+  \lau{TODO: look through this case and find the place where the ''no seal cross boundary thing'' should have been used.
+    (I have done this, and it does not seem necessary (?))}
   In case \ref{case:ftlr:call}, we may assume $r_1 \neq r_2$\footnote{If the register contains a data capability, then the execution fails in the step after the jump.
@@
   \[
-    \npair[n'-1]{(\Phi_S',\Phi_T'')} \in \lro
-  \]
-  By assumption, we have $\npair[n']{(\reg_S(r_i),\reg_T(r_i))} \in \lrrg{\trust}(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
+    \npair[n-1]{(\Phi_S',\Phi_T'')} \in \lro
+  \]
+  By assumption, we have $\npair[n]{(\reg_S(r_i),\reg_T(r_i))} \in \lrrg{\trust}(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
   We know the capabilities in $r_1$ and $r_2$ are sealed capabilities, and by
   Lemma~\ref{lem:unique-h-sigma} and the definition of $\lrvg{\trust}$ we get
-  $\npair[n'-1]{(c_2,c_2')} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$
+  $\npair[n-1]{(c_2,c_2')} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$
   and w.l.o.g.\
@@
   \[
-    \npair[n']{c_1,c_2,c_1',c_2'} \in \lrexj(W_{R,1}'\oplus W_{R,2}')
+    \npair[n-1]{c_1,c_2,c_1',c_2'} \in \lrexj(W_{R,1}'\oplus W_{R,2}')
   \]
@@
   \item \label{case:ftlr:scall:world} $W_{R,1}'\oplus W_{R,2}' \oplus W_R' \oplus W_M'$
-  \item \label{case:ftlr:scall:reg} $\npair[n'-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$
-  \item \label{case:ftlr:scall:mem} $\memSat[n'-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}$
+  \item \label{case:ftlr:scall:reg} $\npair[n-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$
+  \item \label{case:ftlr:scall:mem} $\memSat[n-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}$
   \end{enumerate}
@@
   \[
-    \npair[n'-1]{(\Phi_S',\Phi_T'')} \in \lro
+    \npair[n-1]{(\Phi_S',\Phi_T'')} \in \lro
   \]
@@
     \[
-      \npair[n'-1]{(\stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1},((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1))} \in \lrv(W_{R,\rstk}')
+      \npair[n-1]{(\stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1},((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1))} \in \lrv(W_{R,\rstk}')
     \]
@@
     \[
-      \npair[n']{(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk},((\rw, \linear), \baddr_\stk ,\eaddr_\stk, \aaddr_\stk)} \in \lrvg{\trust}(W_{R,\stk})
+      \npair[n]{(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk},((\rw, \linear), \baddr_\stk ,\eaddr_\stk, \aaddr_\stk)} \in \lrvg{\trust}(W_{R,\stk})
     \]
@@
     \[
-      \npair[n'-1]{[\baddr_\stk,\aaddr_\stk-1]} \in \stackReadCond{W_{R,\rstk}'}
+      \npair[n-1]{[\baddr_\stk,\aaddr_\stk-1]} \in \stackReadCond{W_{R,\rstk}'}
     \]
@@
     \[
-      \npair[n']{[\baddr_\stk,\eaddr_\stk]} \in \stackReadCond{W_{R,\rstk}}
+      \npair[n]{[\baddr_\stk,\eaddr_\stk]} \in \stackReadCond{W_{R,\rstk}}
     \]
@@
     \[
-      \npair[n'-1]{[\baddr_\stk,\aaddr_\stk-1]} \in \stackWriteCond{W_{R,\rstk}'}
+      \npair[n-1]{[\baddr_\stk,\aaddr_\stk-1]} \in \stackWriteCond{W_{R,\rstk}'}
     \]
@@
     \[
-      \npair[n'-1]{(0,0)} \in \lrv(W_{R,\rtmp{1}}')
+      \npair[n-1]{(0,0)} \in \lrv(W_{R,\rtmp{1}}')
     \]
@@
     \[
-      \npair[n'-1]{(\linCons{c_1},\linCons{\Phi_T(r_1)})} \in \lrv(W_{R,r_1}')
+      \npair[n-1]{(\linCons{c_1},\linCons{\Phi_T(r_1)})} \in \lrv(W_{R,r_1}')
     \]
@@
     \[
-      \npair[n'-1]{(\sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)},\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)})} \in \lrv(W_{R,\rretd}')
+      \npair[n-1]{(\sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)},\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)})} \in \lrv(W_{R,\rretd}')
     \]
@@
     \begin{enumerate}[label=\alph*)]
-    \item for $n'' < n'-1$ we have
+    \item for $n'' < n-1$ we have
       \[\npair[n'']{\left(\array{l}\retptrd(\aaddr_\stk,\eaddr_\stk),\\((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)\endarray\right)}
@@
         \begin{array}{l}
-          \forall W'' \future \purePart{W_{R,\rretd}'}, W_o, n'' < n'-1,\\
+          \forall W'' \future \purePart{W_{R,\rretd}'}, W_o, n'' < n-1,\\
           \quad\npair[n'']{(\vsc_S',\vsc_T')}
@@
           \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o) \ldotp \\
-          \qquad\npair[n']{\retptrd(\aaddr_\stk,\eaddr_\stk),\vsc_S',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1),\vsc_T'} \in
+          \qquad\npair[n'']{\retptrd(\aaddr_\stk,\eaddr_\stk),\vsc_S',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1),\vsc_T'} \in
           \lrexj(W'\oplus W_o))
@@
     \[
-      \npair[n'-1]{\left(\array{l}\sealed{\sigma',\retptrc(\baddr,\eaddr,\aaddr +
+      \npair[n-1]{\left(\array{l}\sealed{\sigma',\retptrc(\baddr,\eaddr,\aaddr +
             \calllen)},\\\sealed{\sigma',((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)}\endarray\right)} \in \lrv(W_{R,\rretc}')
@@
     \end{equation}
-    for $n'' < n'-1$.
+    for $n'' < n-1$.
     Both capabilities are non-linear, so
@@
       \begin{array}{l}
-        \forall W'' \future \purePart{W_{R,\rretc}'}, W_o, n'' < n'-1,\\
+        \forall W'' \future \purePart{W_{R,\rretc}'}, W_o, n'' < n-1,\\
         \quad  \npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o) \ldotp \\
@@
         \[
-          \npair[n''']{(((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen))} \in \lreg[\trusted](W''')
+          \npair[n''']{(((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen))} \in \lre[\trusted](W''')
         \]
@@
           We have $\Phi^{(4)}_S.\reg(r) = \reg^{(3)}_S(r)$ and $\Phi^{(5)}_T.\reg(r) = \reg^{(3)}_T(r)$.
-          We already know: $\npair[n''']{(\reg^{(3)}_S(r),\reg^{(3)}_T(r))} \in \lrr(W''_{R,r})$, so by Lemma~\ref{lem:monotonicity}, we get
+          We already know:
+          \[\npair[n''']{(\reg^{(3)}_S(r),\reg^{(3)}_T(r))} \in \lrr(W''_{R,r}),\] so by Lemma~\ref{lem:monotonicity}, we get
           \[

### 2018-06-24 fully-abstract-comp/ftlr-proof-cases.tex
explanation

@@
         \end{enumproof}
-        Assuming the above, we use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$ to get
+        Assuming the above, we use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$ (note that from assumption ``$\Phi_S$ reasonable up to $n$ steps'' and Lemma~\ref{lem:ec-reasonable-downwards-closed} and $n''' \le n'' < n - 1$ we get ``$\Phi_S$ reasonable up to $n'''+1$ steps'' from which it follows that ``$\Phi_S(\pcreg) + \calllen$ reasonable up to $n'''$ steps''.
+        ) To get
         \[

### 2018-06-24 fully-abstract-comp/ftlr-proof-cases.tex
The last of the FTLR call case.

@@
 
-Case: \ref{case:ftlr:scall:mem}, we need to show
-
-\begin{equation}
-  \memSat[n'-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}\label{eq:ftlr:call:memsat-case}
-\end{equation}
-We have the following memory satisfaction on the initial memory:
-\begin{equation}
-  \memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}\label{eq:ftlr-call-og-mem-sat}
-\end{equation}
-By definition of memory satisfaction, this gives us:
-\begin{itemize}
-\item $\stk = (\opc_1,\ms_1) :: \dots :: (\opc_m,\ms_m)$
-\item $\ms_1 \# \dots \# \ms_m \# \ms_S \# \ms_\stk$
-\item $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$
-\item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
-  \begin{itemize}
-  \item $\ms_S = \ms_{f,S} \uplus \ms_S'$
-  \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
-  \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
-  \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
-  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})$
-  \end{itemize}
-\end{itemize}
-We need to consider the memories of $\Phi_S''$ and $\Phi_T''$.
-The memories of the initial configurations have only changed little:
-\[
-  \Phi_T''.\mem = \Phi_T''.\mem\update{\aaddr_\stk}{42}
-\]
-and 
-\[
-\Phi_S''.\ms_\stk = \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
-\]
-and
-\[
-  \Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk
-\]
-with $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
-We need to show the memory satisfaction in \ref{eq:ftlr:call:memsat-case}. Say
-$(\opc_0,\ms_0) = ((\aaddr+\calllen),\ms_{\stk\_\priv,S})$. In this case we have
-to argue 
-\[
-  \ms_{\stk\_\priv,S} \# \ms_1 \# \dots \# \ms_m \# \ms_S \# \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
-\]
-which follows from Assumption~\ref{eq:ftlr:call:memsat-case} and the fact that the memory we remove from $\Phi_S.\ms_\stk$ is exactly what we add to the private stack.
-\lau{This is a bit more involved than I make it to be here, but the fact that we also have a safety assumption on the register-file with the stack pointer means that all of the memory removed from the free stack is actually there.}
-
-  This concludes case \ref{case:ftlr:scall:reg}
-
-  Case: \ref{case:ftlr:scall:mem}, \lau{TODO}
+% Case: \ref{case:ftlr:scall:mem}, we need to show
+
+% \begin{equation}
+%   \memSat[n-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}\label{eq:ftlr:call:memsat-case}
+% \end{equation}
+% We have the following memory satisfaction on the initial memory:
+% \begin{equation}
+%   \memSat[n]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}\label{eq:ftlr-call-og-mem-sat}
+% \end{equation}
+% By definition of memory satisfaction, this gives us:
+% \begin{itemize}
+% \item $\stk = (\opc_1,\ms_1) :: \dots :: (\opc_m,\ms_m)$
+% \item $\ms_1 \# \dots \# \ms_m \# \ms_S \# \ms_\stk$
+% \item $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$
+% \item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
+%   \begin{itemize}
+%   \item $\ms_S = \ms_{f,S} \uplus \ms_S'$
+%   \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
+%   \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
+%   \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
+%   \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})$
+%   \end{itemize}
+% \end{itemize}
+% We need to consider the memories of $\Phi_S''$ and $\Phi_T''$.
+% The memories of the initial configurations have only changed little:
+% \[
+%   \Phi_T''.\mem = \Phi_T''.\mem\update{\aaddr_\stk}{42}
+% \]
+% and 
+% \[
+% \Phi_S''.\ms_\stk = \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
+% \]
+% and
+% \[
+%   \Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk
+% \]
+% with $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
+% We need to show the memory satisfaction in \ref{eq:ftlr:call:memsat-case}. Say
+% $(\opc_0,\ms_0) = ((\aaddr+\calllen),\ms_{\stk\_\priv,S})$. In this case we have
+% to argue 
+% \[
+%   \ms_{\stk\_\priv,S} \# \ms_1 \# \dots \# \ms_m \# \ms_S \# \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
+% \]
+% which follows from Assumption~\ref{eq:ftlr:call:memsat-case} and the fact that the memory we remove from $\Phi_S.\ms_\stk$ is exactly what we add to the private stack.
+% \lau{This is a bit more involved than I make it to be here, but the fact that we also have a safety assumption on the register-file with the stack pointer means that all of the memory removed from the free stack is actually there.}
+
+%   This concludes case \ref{case:ftlr:scall:reg}
+
+  Case \ref{case:ftlr:scall:mem}:
+  we need to show:
+  \[
+    \memSat[n-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}
+  \]
+  which amounts to
+  \[
+    \memSat[n-1]{\Phi_S.\mem,((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S,\Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]},\Phi_T''.\mem}{W_M'}
+  \]
+  for $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
+
+  In order to show this, we will first show the following:
+  \begin{itemize}
+  \item $\baddr_\stk = \stkb$
+
+    We know $\Phi_S$ is reasonable up to $n$ steps.
+    Further, from $\callCond{\Phi_S,r_1,r_2,a}$ and $b \leq a$ and $a + \calllen -1 \leq e$ and $[b,e] \subseteq \ta$ we can conclude that $\Phi_S$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$.
+    By the \emph{Guarantee stack base address before call} we then know $\Phi_S(r_\stk) = \stkptr{\_,\stkb,\_,\_}$.
+  \end{itemize}
+
+  By assumption we have
+  $\memSat{\Phi_S.\mem,\Phi_S.\stk,\Phi_S.\ms_\stk,\Phi_T.\mem}{W_M}$ which
+  gives us $\ms_\var{T,stack}, \ms_\var{T,free\_stack},$ $\ms_\var{T,heap},
+  \ms_{T,f}, \ms_{S,f}, \ms_S', W_{M,\var{stack}}, W_{M,\var{free\_stack}},
+  W_{M,\var{heap}}$ such that:
+  \begin{itemize}
+  \item $\Phi_S.\stk = (\opc_0, \ms_0) \dots (\opc_m,\ms_m)$
+  \item $\ms_0 \uplus \dots \uplus \ms_m \uplus \Phi_S.\mem \uplus \Phi_S.\ms_\stk$
+  \item $W_M = W_{M,\var{stack}} \oplus W_{M,\var{free\_stack}} \oplus W_{M,\var{heap}}$
+  \item $\Phi_S.\mem = \ms_{f,S} \uplus \ms_S'$
+  \item $\Phi_T.\mem = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus
+    \ms_\var{T,heap} \uplus \ms_{T,f}$
+  \item $\memSatStack{\Phi_S.\stk,\ms_\var{T,stack}}{W_{M,\var{stack}}}$
+  \item $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}}$
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{M,\var{heap}})$
+  \end{itemize}
+
+  In order to prove the memory satisfaction, we pick the same memories except for the following changes:
+  \begin{itemize}
+  \item The target free stack partition: $\ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}$
+  \item The target private stack partition: $\ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$
+  \end{itemize}
+  and the worlds
+  \begin{itemize}
+  \item Free stack world: $W_{M,\var{free\_stack}}'$ is $W_{R,1}'$ with the ownership of $W_{M,\var{free\_stack}}$ except that it gives up any ownership that we used for the safety of the addresses $[\aaddr_\stk,\eaddr_\stk]$.
+  \item Private stack world: $W_{M,\var{stack}}'$ is $W_{R,1}'$ with the ownership of $W_{M,\var{stack}}$ except that it takes the ownership that $W_{M,\var{free\_stack}}$ used for the safety of the addresses $[\aaddr_\stk,\eaddr_\stk]$.
+  \item Heap world: $W_{M,\var{heap}}'$ is $W_{R,1}'$ with the ownership of $W_{M,\var{heap}}$ which gives us $W_{M,\var{heap}}' \future W_{M,\var{heap}}$.
+  \end{itemize}
+  We now need to show:
+  \begin{itemize}
+  \item $\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: (\opc_0, \ms_0) \dots (\opc_m,\ms_m)$
+
+    Trivial.
+  \item $\ms_{\stk\_\priv,S} \uplus \ms_0 \uplus \dots \uplus \ms_m \uplus \Phi_S.\mem \uplus \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}$
+
+    Follows by assumption and the fact that we have shuffled around some memory\footnote{Thanks to the register-file safety assumption we know for sure that the memory we remove from the free stack is actually there.}.
+  \item $W_M' = W_{M,\var{stack}}' \oplus W_{M,\var{free\_stack}}' \oplus W_{M,\var{heap}}'$
+
+    This follows by assumption and the fact that all ownership we have added to a world has been removed from another.
+  \item $\Phi_S.\mem = \ms_{f,S} \uplus \ms_S'$
+
+    By assumption.
+  \item $\Phi_T.\mem = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} \uplus \ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
+
+    Follows by assumption and the fact that we have shuffled around some memory\footnote{Thanks to the register-file safety assumption we know for sure that the memory we remove from the free stack is actually there.}.
+  \item $\memSatStack{((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S,\ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}}{W_{M,\var{stack}}'}$:
+
+    For most of the conditions, they follow from the previous stack satisfaction assumption.
+    The only challenge is to argue that the new stack frame satisfies the conditions.
+    First of all, we know $\ms_{\stk\_\priv,S}$ is non-empty as it at least contains address $\aaddr_\stk$.
+    Next, we need to argue that
+    \[
+      \begin{array}{l}
+        \forall i \in \{0,\dots,m\} \ldotp \\
+        \quad \forall a \in \dom(\ms_{\stk\_\priv,S}) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
+      \end{array}
+    \]
+    The first bit, $\stkb < a$ for $a \in \dom(\ms_{\stk\_\priv,S})$ follows from the fact that $a_\stk$ is the smallest address of $a \in \dom(\ms_{\stk\_\priv,S})$ and by assumption $\stkb < a_\stk$.
+
+    Now assume for contradiction that there exists $a' \in \dom(\ms_j)$ for some $j$ such that $a' \le a$ for some $a \in \dom(\ms_{\stk\_\priv,S})$.
+    By assumption we have $\stkb < a'$ so $a' \in [\stkb+1,a] \subseteq [\stkb+1,a_\stk] \subseteq [\stkb,e_\stk]$ which means that $a'$ is an address governed by a stack pointer.
+    By the register-safety assumption this must mean that it is an address of the free part of the stack.
+    At the same time, it must be an address of the private stack because $\ms_j$ is part of the stack from the original configuration.
+    This contradicts the initial memory satisfaction assumption as the different parts must be disjointed.
+
+    From the stack satisfaction assumption, we get $R_\ms$ and $R_W$. Pick
+    \[
+      R_\ms'(r) =
+      \begin{cases}
+        (\ms_{\stk\_\priv,S}, (\aaddr+\calllen),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42})& \text{for $r = r_\var{priv\_stk}$}\\
+        R_\ms(r) & \totherwise
+      \end{cases}
+    \]
+    and for $R_W'$ pick
+    \[
+      R_W'(r) =
+      \begin{cases}
+        W_{M,\var{free\_stack},[\aaddr_\stk,\eaddr_\stk]}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)]
+        & \text{for $r = r_\var{priv\_stk}$} \\
+        R_W(r)[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)] &
+      \end{cases}
+    \]
+    where $W_{M,\var{free\_stack},[\aaddr_\stk,\eaddr_\stk]}$ is the world that the free stack assumption uses to satisfy that range of addresses.
+
+    Most of the condition trivially holds. 
+    % This follows in part by mono of the interpretation function and the worlds
+    % picked are future worlds of the past worlds.
+    The only one that requires some argumentation is
+    \[
+      \npair[n-1]{(\ms_{\stk\_\priv,S},\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42})}
+      \in
+      \stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square}.H
+      \; \xi^{-1}(R_W'(r_\var{priv\_stk}))
+    \]
+    Where $R_W'(r_\var{priv\_stk})$ is the part of $W_{M,\var{stack}}'$ with the ownership used for addresses $[\aaddr_\stk, \eaddr_\stk]$ in the memory satisfaction assumption.
+
+    As we have the safe register-file assumption, we know that the stack capability is safe.
+    This means that addresses $[\aaddr_\stk,\eaddr_\stk]$ must be part of the free stack.
+    Further, $\Phi''_T.\mem = \Phi_T.\mem\update{\aaddr_\stk}{42}$ and $\ms_\var{T,free\_stack} \subseteq \Phi_T.\mem$ and $\dom(\ms_\var{T,free\_stack}) \supseteq [\aaddr_\stk,\eaddr_\stk]$ from which it follows that the memories are equal to the one of the static region.
+
+    It remains to show
+    \[
+      \forall \aaddr \in [\aaddr_\stk,\eaddr_\stk] \ldotp \npair[n-1]{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
+    \]
+    for $\aaddr = \aaddr_\stk$ it is trivial as we have to show
+    \[
+      \npair[n-1]{(42,42)} \in \lrv(W_{R,\aaddr_\stk}')
+    \]
+    for $\aaddr \in [\aaddr_\stk+1,\eaddr_\stk]$ we need to show
+    \[
+      \npair[n-1]{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
+    \]
+    If we can show
+    \[
+      \npair[n-1]{(\Phi_S.\ms_\stk(\aaddr),\ms_\var{T,free\_stack}(\aaddr))} \in \lrv(W_{R,\aaddr})
+    \]
+    then we are done by monotonicity of $\lrv$.
+
+    By assumption we know $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$ which entails $\npair{(\Phi_S.\reg(\rstk),\Phi_T.\reg(\rstk))} \in \lrrg{\trust}(W_{R,\rstk})$.
+    We know $\Phi_S.\reg(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk}$, so by the definition of $\lrv$ we get
+    \[
+      \npair{[\stkb,\eaddr_\stk]} \in \stackReadCond{W_{R,\rstk}}
+    \]
+    which in turn gives us $S_{\rstk} \subseteq \addressable{\linear,\pwfree}$ and
+    $R_{\rstk} : S_{\rstk} \fun \powerset{\nats}$ for which
+    \begin{itemize}
+    \item $\forall r \in S_{\rstk} \ldotp |R_{\rstk}(r)| = 1$
+    \item $\uplus_{r \in S_{\rstk}} R_{\rstk}(r) \supseteq [\stkb,\eaddr_\stk]$
+    \item $\forall r \in S_{\rstk} \ldotp W_{R,\rstk}(r).H \nsubeq \stdreg{R_{\rstk}(r),\gc}{\spao}$
+    \end{itemize}
+    Further, we know $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}.\free}$ which means that we have $R_\ms : \dom(\activeReg{W_{M,\var{free\_stack}}}) \fun MemSeg \times \MemSeg$ and $R_W : \dom(W_{M,\var{free\_stack}}.\free) \fun \Worlds$.
+    $R_W$ distributes the ownership of $W_{M,\var{free\_stack}}$ and $R_\ms$ partitions the memories.
+
+    We know that all regions in $S_{\rstk}$ govern singleton memory segments, so $R_\ms$ must map to singleton memory segment pairs for $r \in S_{\rstk}$.
+    Further, by definition of the free stack satisfaction for $r \in \S_{\rstk}$ we have
+    \[
+      \npair{R_\ms(r)} \in W_{R,\rstk}(r).H \; \xi^{-1}(R_W(r))
+    \]
+    which entails
+    \[
+      \npair[n-1]{R_\ms(r)} \in \stdreg{R_{\rstk}(r),\gc}{\spao} \; \xi^{-1}(R_W(r))
+    \]
+    which entails
+    \[
+      \npair[n-1]{R_\ms(r)(a)} \in \lrv (R_W(r))
+    \]
+    for $a \in R_{\rstk(r)}$.
+
+    Now this is exactly what we wanted to show because $R_W(r)$ is what we picked as $W_{R,\aaddr}$.
+  \item $\memSatFStack{\Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]},\ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}}{W_{M,\var{free\_stack}}'}$:
+
+    From the safety assumption on the stack capability, we can deduce a number of things:
+    \begin{itemize}
+    \item $[\aaddr_\stk,\eaddr_\stk]$ must have been part of the free stack
+    \item for every address in $[\aaddr_\stk,\eaddr_\stk]$ there is a region for that singleton memory segment.
+    \end{itemize}
+    The first part means that we do indeed remove all of the memory we try to subtract.
+    The latter means that we can reuse the same split of the remaining memory and the world ownership as we get from assumption $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}}$.
+    Using this, the result follows from monotonicity of the $H$ function.
+
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{M,\var{heap}}')$:
+
+    Follows from Lemma~\ref{lem:heap-sat-unchanged-future}, the fact the the heap part of the world remains unchanged and $W_{M,\var{heap}}' \future W_{M,\var{heap}}$.
+  \end{itemize}
 \end{proof}
@@
 \end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:heap-sat-unchanged-future}
+  If $\pwheap = \pwheap[W']$ and $W' \future W$ and $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap(W)$, then $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in\lrheap(W')$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
 

### 2018-06-24 fully-abstract-comp/ftlr-proof-cases.tex
More of the FTLR proof and other changes.
Sorry about the big commit.

@@
   The result then follows easily by definition of $\lrol$ and $\lror$.
+\end{proof}
+
+\begin{lemma}[Capability safety doesn't depend on address]
+  \label{lem:cap-in-lrv-regardless-of-addr}
+  If
+  \[
+    \npair{((\perm,\lin),\baddr,\eaddr,\aaddr),((\perm,\lin),\baddr,\eaddr,\aaddr)} \in \lrvg{\trust}(W)
+  \]
+  then
+  \[
+    \npair{((\perm,\lin),\baddr,\eaddr,\aaddr'),((\perm,\lin),\baddr,\eaddr,\aaddr')} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  TODO (Should be trivial)
+\end{proof}
+
+\begin{lemma}[Stack capability safety doesn't depend on address]
+\label{lem:stkptr-in-lrv-regardless-of-addr}
+  If
+  \[
+    \npair{(((\perm,\lin),\baddr,\eaddr,\aaddr),\stkptr{\perm,\baddr,\eaddr,\aaddr})} \in \lrvg{\trust}(W)
+  \]
+  then
+  \[
+    \npair{(((\perm,\lin),\baddr,\eaddr,\aaddr'),\stkptr{\perm,\baddr,\eaddr,\aaddr'})} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  TODO (Should be trivial)
+\end{proof}
+
+\begin{lemma}[Seal safety doesn't depend on current seal]
+  \label{lem:seal-in-lrv-regardless-of-cur-seal}
+  If
+  \[
+    \npair{(\seal{\sigma_\baddr,\sigma_\eaddr,\sigma},\seal{\sigma_\baddr,\sigma_\eaddr,\sigma})} \in \lrvg{\trust}(W)
+  \]
+  then
+  \[
+    \npair{(\seal{\sigma_\baddr,\sigma_\eaddr,\sigma'},\seal{\sigma_\baddr,\sigma_\eaddr,\sigma'})} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  TODO (Should be trivial)
+\end{proof}
+
+\begin{lemma}[Capability safety monotone w.r.t. permission]
+  \label{lem:cap-in-lrv-mono-perm}
+  If
+  \[
+    \npair{((\perm,\lin),\baddr,\eaddr,\aaddr),((\perm,\lin),\baddr,\eaddr,\aaddr)} \in \lrvg{\trust}(W)
+  \]
+  and
+  \[
+    \perm' \sqsubseteq \perm
+  \]
+  then
+  \[
+    \npair{((\perm',\lin),\baddr,\eaddr,\aaddr),((\perm',\lin),\baddr,\eaddr,\aaddr)} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  TODO (Should be trivial)
+\end{proof}
+
+\begin{lemma}[Stack capability safety monotone w.r.t. permission]
+\label{lem:stkptr-in-lrv-mono-perm}
+  If
+  \[
+    \npair{(((\perm,\lin),\baddr,\eaddr,\aaddr),\stkptr{\perm,\baddr,\eaddr,\aaddr})} \in \lrvg{\trust}(W)
+  \]
+  and
+  \[
+    \perm' \sqsubseteq \perm
+  \]
+  then
+  \[
+    \npair{(((\perm',\lin),\baddr,\eaddr,\aaddr),\stkptr{\perm',\baddr,\eaddr,\aaddr})} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  TODO (Should be trivial)
 \end{proof}
@@
     \end{itemize}
-  \item All of the following hold: (includes cap-manipulation cases: move, cca,
+  \item \label{case:ftlr:move-cca-etc} All of the following hold: (includes cap-manipulation cases: move, cca,
     restrict, seta2b, cseal, split, splice, that can be handled mostly
@@
       \begin{enumproof}
-      \item (restrict,cca,seta2b) $w_1 = ((\perm',\lin),\baddr,\eaddr,\aaddr')$,
+      \item \label{case:ftlr:move-cca-etc:cca-normal} (restrict,cca,seta2b) $w_1 = ((\perm',\lin),\baddr,\eaddr,\aaddr')$,
         $w_1' = ((\perm',\lin),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) =
@@
         1$
-      \item (restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$,
+      \item \label{case:ftlr:move-cca-etc:cca-stack}(restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$,
         $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) =
@@
         $k = 1$
-      \item (cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$,
+      \item \label{case:ftlr:move-cca-etc:cca-seal}(cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$,
         $w_1' = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $\Phi_S(r_1) =
-        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$ and $\Phi_T(r_1) =
-        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$, $k = 1$
-      \item (move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S(r_1)$, and
-        $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg \isLinear{w_1}
-        \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$) and $k = 2$.
-      \item (cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' =
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\Phi_T(r_1) =
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k = 1$
+      \item \label{case:ftlr:move-cca-etc:move}(move) $w_1 = \Phi_S(r_2)$, $w_1' = \Phi_S'(r_2)$, and $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg \isLinear{w_1} \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$) and $r_1 \neq r_2$ and $k = 2$.
+      \item \label{case:ftlr:move-cca-etc:move-eq}(move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S'(r_1)$ and $k = 1$.
+      \item \label{case:ftlr:move-cca-etc:cseal}(cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' =
         \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) =
         \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k=1$
-      \item (split) $\Phi_T(r_3) = \Phi_S(r_3) =
+      \item \label{case:ftlr:move-cca-etc:split-normal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
         ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$,
@@
         ((\perm,\lin),n+1,\eaddr,\aaddr)$, $k=2$.
-      \item (split) $\Phi_T(r_3) = \Phi_S(r_3) =
+      \item \label{case:ftlr:move-cca-etc:split-seal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
         \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \leq n$, $n <
@@
         w_2' = \seal{n+1,\sigma_\eaddr,\sigma}$, $k=2$.
-      \item (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$,
+      \item \label{case:ftlr:move-cca-etc:split-stack} (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$,
         $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$,
@@
         ((\perm,\linear),n+1,\eaddr,\aaddr)$, $k=2$.
-      \item (splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
+      \item \label{case:ftlr:move-cca-etc:splice-stack}(splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
         $\Phi_S(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_}$, and $\Phi_T(r_3) =
@@
         = 0$, and $k=3$
-      \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
+      \item \label{case:ftlr:move-cca-etc:splice-normal}(splice) $\Phi_T(r_2) = \Phi_S(r_2) =
         ((\perm,\lin),\baddr_2,\eaddr_2,\_)$, and $\Phi_T(r_3) = \Phi_S(r_3) =
@@
         \Phi_T(r_3))$, and $k=3$
-      \item (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
+      \item \label{case:ftlr:move-cca-etc:splice-seal} (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
         \seal{\sigma_{\baddr,2},\sigma_{\eaddr,2},\_}$, and $\Phi_T(r_3) =
@@
     \end{itemize}
-  \item All of the following hold: (includes memory-manipulation cases: store,
+  \item \label{case:ftlr:store-load-mem} All of the following hold: (includes memory-manipulation cases: store,
     load, that can be handled mostly uniformly)
@@
     \end{itemize}
-  \item All of the following hold: (includes memory-manipulation cases: store,
+  \item \label{case:ftlr:store-load-stack} All of the following hold: (includes memory-manipulation cases: store,
     load, that can be handled mostly uniformly (stack))
@@
     \end{itemize}
-  \item All of the following hold: (includes control-flow manipulation cases:
+  \item \label{case:ftlr:store-load-jump} All of the following hold: (includes control-flow manipulation cases:
     jmp, jnz, xjmp, that can be handled mostly uniformly)
@@
   \end{itemize}
+
+  Case \ref{case:ftlr:move-cca-etc}:
+
+  First show: if $\lin = \normal$, then
+  \[
+    \npair[n-1]{(((\rx,\normal),\baddr,\eaddr,\aaddr),((\rx,\normal),\baddr,\eaddr,\aaddr))} \in \lrvg{\trust}(\purePart{W_R})
+  \]
+  First observe that by Lemma~\ref{lem:purePart-oplus} $\purePart{W_R} = \purePart{W_\pcreg}$.
+  Further, by assumption, Lemma~\ref{lem:downwards-closed}, Lemma~\ref{lem:purePart-oplus}, and Lemma~\ref{lem:non-linear-pure}, we have
+  \[
+    \npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{}(\purePart{W_R})
+  \]
+
+  If $\trust = \trusted$, then by assumption we have $[\baddr,\eaddr] \subseteq \ta$ which means that all the conditions for are met for the capability pair to be in the trusted part of the value relation.
+
+  If $\trust = \untrusted$, then we need to show that the capability pair is in the untrusted part of the value relation which means that we need to show:
+  \begin{itemize}
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{}(\normal,\purePart{W_R})$
+
+    This follows by assumption, Lemma~\ref{lem:non-linear-pure}, Lemma~\ref{lem:purePart-oplus}, and Lemma~\ref{lem:downwards-closed}.
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{}(\purePart{W_R})$
+
+    We already showed this.
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \execCond{}(\purePart{W_R})$
+
+    To this end let $W' \future \purePart{W_R}$ and $n' < n-1$ and $\aaddr' \in [\baddr',\eaddr']$ be given, and show
+    \[
+      \npair[n']{(((\rx,\normal),\baddr',\eaddr',\aaddr'),((\rx,\normal),\baddr',\eaddr',\aaddr'))} \in \lre(W')
+    \]
+    If $[\baddr',\eaddr'] = \emptyset$, then this is trivial, so assume $[\baddr',\eaddr'] \neq \emptyset$.
+    Further, for $n'' \leq n'$ assume
+    \begin{itemize}
+    \item $\npair[n'']{\reg_S',\reg_T'} \in \lrr(W_R')$
+    \item $\memSat{\ms_S',\stk',\ms_\stk',\ms_T'}{W_M'}$
+    \item $W' \oplus W_R' \oplus W_M'$ is defined
+    \end{itemize}
+    and show
+    \[
+      \npair[n'']{\left(
+          \begin{array}{l}
+          (\ms_S',\reg_S'\update{\pcreg}{((\rx,\normal),\baddr',\eaddr',\aaddr')},\stk',\ms_\stk') \\
+          (\ms_T',\reg_T'\update{\pcreg}{((\rx,\normal),\baddr',\eaddr',\aaddr')})
+          \end{array}
+         \right)} \in \lro
+    \]
+    To this end, we want to use the IH which means that we need the following:
+    \begin{itemize}
+    \item $\trust = \untrusted$ and $\npair[n'']{[\baddr',\eaddr']} \in \readCond{W'}$
+
+      This follows from the readCondition assumption, Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:downwards-closed}, Lemma~\ref{lem:monotonicity}, Lemma~\ref{lem:purePart-oplus}, and Lemma~\ref{lem:non-linear-pure}.
+    \item $\npair[n'']{[\baddr',\eaddr']} \in \xReadCond{W'}$
+
+      This follows from what we already showed about the readXCondition,
+      Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:downwards-closed}, and Lemma~\ref{lem:monotonicity}
+    \item $\npair[n'']{\reg_S',\reg_T'} \in \lrr(W_R')$
+
+      By assumption.
+    \item $\memSat{\ms_S',\stk',\ms_\stk',\ms_T'}{W_M'}$
+
+      By assumption.
+
+    \item $W' \oplus W_R' \oplus W_M'$ is defined
+
+      By assumption
+    \end{itemize}
+  \end{itemize}
+
+  Now show
+  \[
+    \npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R)
+  \]
+  assuming that neither of the following is the case:
+  \begin{itemize}
+  \item Case \ref{case:ftlr:move-cca-etc:move} and $\isLinear{w_2}$ and $r_2 = \pcreg$.
+  \item Case \ref{case:ftlr:move-cca-etc:cseal} and $\isLinear{\Phi_S(r_1)}$ and $r_1 = \pcreg$.
+  \end{itemize}
+
+  By assumption we have
+  \[
+    \npair[n]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)
+  \]
+  which gives us $R_R : (\RegName \setminus \{\pcreg\}) \fun World$ such that $W_R = \bigoplus_{r\in \RegName \setminus \{\pcreg\}} R_R(r)$ and for all $r$ in $\RegName \setminus \{\pcreg\}$ we have $\npair[n]{(\Phi_S.\reg(r),\Phi_T.\reg(r))} \in \lrrg{\trust}(R_R(r))$.
+  To this end, we need to consider each of cases \ref{case:ftlr:move-cca-etc:cca-normal}-\ref{case:ftlr:move-cca-etc:splice-seal}:
+  \begin{itemize}
+  \item Case \ref{case:ftlr:move-cca-etc:cca-normal}:
+
+    Pick $R_R$ as the ownership distribution.
+    For $r \neq r_1$ it follows by assumption and Lemma~\ref{lem:downwards-closed}.
+    For $r = r_1$ it also follows by assumption and Lemma~\ref{lem:downwards-closed}, \ref{lem:cap-in-lrv-regardless-of-addr}, and \ref{lem:cap-in-lrv-mono-perm}.
+  \item Case \ref{case:ftlr:move-cca-etc:cca-stack}:
+
+    Pick $R_R$ as the ownership distribution.
+    For $r \neq r_1$ it follows by assumption and Lemma~\ref{lem:downwards-closed}.
+    For $r = r_1$ it also follows by assumption and Lemma~\ref{lem:downwards-closed}, \ref{lem:stkptr-in-lrv-regardless-of-addr}, and \ref{lem:stkptr-in-lrv-mono-perm}.
+  \item Case \ref{case:ftlr:move-cca-etc:cca-seal}:
+
+    Pick $R_R$ as the ownership distribution.
+    For $r \neq r_1$ it follows by assumption and Lemma~\ref{lem:downwards-closed}.
+    For $r = r_1$ it also follows by assumption and Lemma~\ref{lem:downwards-closed} and \ref{lem:seal-in-lrv-regardless-of-cur-seal}.
+
+  \item Case \ref{case:ftlr:move-cca-etc:move}:
+
+    Pick the ownership distribution based on the linearity of $w_2$: If $\isLinear{w_2}$, then pick
+    \[
+      R_R'(r) =
+      \begin{cases}
+        R_R(r_2) \oplus R(r_1) & r = r_1 \\
+        \purePart{W_R}         & r = r_2 \\
+        R_R(r)                 & \totherwise
+      \end{cases}
+    \]
+    if $\neg \isLinear{w_2}$, then pick
+    \[
+      R_R'(r) = R_R(r)
+    \]
+
+    In the case where $\isLinear{w_2}$, we may assume $r_2 \neq \pcreg$. We need to show (assuming $r_1 \neq \pcreg$)
+    \[
+      \npair[n-1]{(\Phi_S'(r_1),\Phi_T'(r_1))} \in \lrvg{\trust}(R_R(r_2) \oplus R(r_1))
+    \]
+    which is 
+    \[
+      \npair[n-1]{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(R_R(r_2) \oplus R(r_1))
+    \]
+    this follows by assumption and Lemma~\ref{lem:monotonicity}.
+
+    We also need to show
+    \[
+      \npair[n-1]{(\Phi_S'(r_2),\Phi_T'(r_2))} \in \lrvg{\trust}(\purePart{W_R})
+    \]
+    which is trivial as $\Phi_S'(r_2) = \Phi_T'(r_2) = 0$.
+
+    Finally for $r \neq r_1,r_2,\pcreg$ 
+    \[
+      \npair[n-1]{(\Phi_S'(r),\Phi_T'(r))} \in \lrvg{\trust}(R_R(r))
+    \]
+    Follows by assumption and Lemma~\ref{lem:downwards-closed}.
+
+    In the case where $\neg \isLinear{w_2}$
+
+    If $r_2 \neq \pcreg$, then for $r \neq r_1$
+    \[
+      \npair[n-1]{(\Phi_S'(r),\Phi_T'(r))} \in \lrvg{\trust}(R_R(r))
+    \]
+    Follows by assumption and Lemma~\ref{lem:downwards-closed}.
+
+    For $r \neq r_1,r_2$
+    \[
+      \npair[n-1]{(\Phi_S'(r_1),\Phi_T'(r_1))} \in \lrvg{\trust}(R_R(r_1))
+    \]
+    amounts to
+    \[
+      \npair[n-1]{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(R_R(r_1))
+    \]
+    By assumption and Lemma~\ref{lem:purePart-oplus} and \ref{lem:non-linear-pure}.
+
+    If $r_2 = \pcreg$, show
+    \[
+      \npair[n-1]{(\Phi_S(\pcreg),\Phi_T(\pcreg))} \in \lrvg{\trust}(R(r_1))
+    \]
+    which follows from Lemma~\ref{lem:monotonicity} and \ref{lem:world-fut-purePart} and what we have proven about the pc.
+    \lau{remember to change this if we do not manage to prove that the pc is safe.}
+  \item Case \ref{case:ftlr:move-cca-etc:move-eq}:
+
+    Follows from Assumption and Lemma~\ref{lem:downwards-closed}.
+
+  \item Case \ref{case:ftlr:move-cca-etc:cseal}:
+
+TODO
+  \item Case \ref{case:ftlr:move-cca-etc:split-normal}:
+
+TODO
+  \item Case \ref{case:ftlr:move-cca-etc:split-seal}:
+
+TODO
+  \item Case \ref{case:ftlr:move-cca-etc:split-stack}:
+
+TODO
+  \item Case \ref{case:ftlr:move-cca-etc:splice-stack}:
+
+TODO
+  \item Case \ref{case:ftlr:move-cca-etc:splice-normal}:
+
+TODO
+  \item Case \label{case:ftlr:move-cca-etc:splice-seal}:
+
+TODO
+  \end{itemize}
+
+  By Lemma~\ref{lem:lro-anti-red-gen} it suffices to show
+  \[
+    \npair[n-1]{\Phi_S',\Phi_T'} \in \lro
+  \]
+  which follows from the induction hypothesis in the case where $\Phi_S'.\reg(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$.
+  In this case, the IH is applicable because we have the following:
+  \begin{itemize}
+  \item One of the following sets of requirements holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S'$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \end{itemize}
+
+    We know one of the following holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \end{itemize}
+    If the latter is the case, then the result follows by Lemma~\ref{lem:downwards-closed}.
+
+    If the former holds, then it follows from Lemma~\ref{lem:ec-reasonable-antired}.
+
+    \item TODO (the remaining cases)
+  \end{itemize}
+
+  It remains to show the case where $\Phi_S'.\reg(\pcreg) \neq ((\rx,\lin),\baddr,\eaddr,\_)$
+
+  We still need to show the following cases:
+  \begin{itemize}
+  \item Case \ref{case:ftlr:move-cca-etc:move} and $\isLinear{w_2}$ and$r_2 = \pcreg$.
+    TODO
+  \end{itemize}
+
+  Case \ref{case:ftlr:store-load-mem}
+
+  TODO
+
+  Case \ref{case:ftlr:store-load-stack}
+
+  TODO
+
+  Case \ref{case:ftlr:store-load-jump}
+
+  TODO
 \end{proof}

### 2018-06-24 fully-abstract-comp/technical-report.tex
typos and such
More of the FTLR proof and other changes.
Sorry about the big commit.

@@
       \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
-      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
+      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
       \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
@@
 \dominique{22-6-2018: address-stratified is useless for length-one regions.}
-
 Note: this new version of execCond uses the expression relation for regular jumps since it expresses the validity of jumping to an address pointed to by an executable capability. 
@@
   \label{def:reasonable-conf}
-  We say that an execution configuration $\Phi$ is reasonable up to $n$ steps
+  We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb)$ 
   iff for all $n' \le n$:
@@
     \item $\src{\Phi'}(\src{r_\stk}) =
-      \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
+      \src{\stkptr{\_,\stkb,\_,\_}}$
     \item $r_1 \neq \rtmp{1}$
@@
     $\nonLinear{w_2}$), then
-    $\npair{(w_1,w_2)} \in \lrv(H_\sigma,\purePart{W})$.
+    $\npair{(w_1,w_2)} \in \lrv(\purePart{W})$.
   \end{itemize}
@@
       \end{equation*}
-      together with \ref{lem:monotonicity} and the fact that $R_W(\addr) \future \purePart(W')$ (by Lemma~\ref{lem:auth-partial-betterthannothing}).
+      together with \ref{lem:monotonicity} and the fact that $R_W(\addr) \future \purePart{W'}$ (by Lemma~\ref{lem:auth-partial-betterthannothing}).
 

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Prove downwards closure lemma.

@@
   \begin{itemize}
+  \item If $\npair{\lin,\baddr,\eaddr} \in \lrp(W)$, then $\npair[n']{\lin,\baddr,\eaddr} \in \lrp(W)$.
   \item If $\npair{A} \in \readCond{\lin,W}$, then $\npair[n']{A} \in \readCond{\lin,W}$.
@@
   \item If $\npair{(w,w')} \in \lrrg{\trust}(W)$, then $\npair[n']{(w,w')} \in \lrrg{\trust}(W)$.
+  \item If $\memSatFStack{ms_\stk,\ms_T}{W}$, then $\memSatFStack[n']{ms_\stk,\ms_T}{W}$.
+  \item If $\memSatStack{ms_\stk,\ms_T}{W}$, then $\memSatStack[n']{ms_\stk,\ms_T}{W}$.
+  \item If $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap{\pwheap}(W)$, then $\npair[n']{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap{\pwheap}(W)$.
   \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$, then also $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$.
@@
 \end{lemma}
+\begin{proof}
+  The properties for $\lrp$, $\readCond{}$, $\stackReadCond{}$, $\writeCond{}$, $\stackWriteCond{}$, $\execCond{}$, $\xReadCond{}$ follow easily by definition.
+  The property for $\lrv$ follows from the others and by definition.
+  The property for $\lrr$ follows directly from the one for $\lrv$.
+
+  The property for $\lrheap$, $\memSatFStack{\_,\_}{\_}$, $\memSatStack{\_,\_}{\_}$ follows from the downward closure of regions in the world, and the property for $(\_,\_,\_,\_) :_n W$ follows from those.
+\end{proof}
 

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Remove outdated comment.
Check some easy lemmas.

@@
 
-\dominique{9-3-2018: the below is a proposed generalisation of the above antireduction lemma  that I expect will hold (based on experience with similar LRs in the past).}
 \begin{lemma}[$\lro$ closed under antireduction (generalised previous lemma)]
@@
 \begin{proof}
-  TODO (Should be trivial)
+  Direct form the definition of $\lrvg{\trust}$.
 \end{proof}
@@
   \[
-    \npair{(((\perm,\lin),\baddr,\eaddr,\aaddr),\stkptr{\perm,\baddr,\eaddr,\aaddr})} \in \lrvg{\trust}(W)
+    \npair{((\stkptr{\perm,\baddr,\eaddr,\aaddr},(\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrvg{\trust}(W)
   \]
@@
   \[
-    \npair{(((\perm,\lin),\baddr,\eaddr,\aaddr'),\stkptr{\perm,\baddr,\eaddr,\aaddr'})} \in \lrvg{\trust}(W)
-  \]
-\end{lemma}
-\begin{proof}
-  TODO (Should be trivial)
+    \npair{(\stkptr{\perm,\baddr,\eaddr,\aaddr'},((\perm,\lin),\baddr,\eaddr,\aaddr'))} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  Direct from the definition of $\lrvg{\trust}$.
 \end{proof}
@@
 \begin{proof}
-  TODO (Should be trivial)
+  Direct from the definition of $\lrvg{\trust}$.
 \end{proof}
@@
 \begin{proof}
-  TODO (Should be trivial)
+  Direct from the definition of $\sqsubseteq$ and $\lrvg{\trust}$, observing that $\npair{[b,e]} \in \readCond{\lin,W}$ implies $\npair{(\lin,b,e)}\in\lrp(W)$ by definition.
 \end{proof}
@@
   \[
-    \npair{(((\perm,\lin),\baddr,\eaddr,\aaddr),\stkptr{\perm,\baddr,\eaddr,\aaddr})} \in \lrvg{\trust}(W)
+    \npair{(\stkptr{\perm,\baddr,\eaddr,\aaddr},((\perm,\lin),\baddr,\eaddr,\aaddr))} \in \lrvg{\trust}(W)
   \]
@@
   \[
-    \npair{(((\perm',\lin),\baddr,\eaddr,\aaddr),\stkptr{\perm',\baddr,\eaddr,\aaddr})} \in \lrvg{\trust}(W)
-  \]
-\end{lemma}
-\begin{proof}
-  TODO (Should be trivial)
+    \npair{(\stkptr{\perm',\baddr,\eaddr,\aaddr},((\perm',\lin),\baddr,\eaddr,\aaddr))} \in \lrvg{\trust}(W)
+  \]
+\end{lemma}
+\begin{proof}
+  \dominique{Doesn't hold: lrp doesn't make sense for stack regions!}
+  TODO
 \end{proof}

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Make callCondition mention offpc and offsigma.

@@
     \begin{itemize}
-    \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
+    \item $\callCond{\Phi_S,r_1,r_2,\offpc,\offsigma,\aaddr}$
     \item $r_1 \neq \rtmp{1}$ and $r_2 \neq \rtmp{1}$
@@
     \item $\baddr \leq \aaddr+\offpc \leq \eaddr$
-      \lau{TODO: Need to have an explicit $\offpc$.}
     \item $\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}$
@@
 
-    \item $\Phi_S' =\var{xjumpResult}\left(c_1,c_2, \Phi_S' \right)$
+    \item $\Phi_S' =\var{xjumpResult}\left(c_1,c_2, \Phi_S'' \right)$
 
@@
       We immediately get $\dom(\ms_{\stk\_\priv,S}) = \dom(\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}) = [\aaddr_\stk,\eaddr_\stk]$.
-      Next, $\decInstr{\mscode'(\aaddr,\aaddr+\calllen-1)} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ follows from $\callCond{\Phi_S,r_1,r_2,\aaddr}$.
+      Next, $\decInstr{\mscode'(\aaddr,\aaddr+\calllen-1)} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ follows from $\callCond{\Phi_S,r_1,r_2,\offpc,\offsigma,\aaddr}$.
       Finally, $\ms_\code(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr}$ with $\sigma' = \sigma_\baddr + \offpc \in \sigrets'$ which follows from $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$ and the fact that the call is there.
@@
     We know $\Phi_S$ is reasonable up to $n$ steps.
-    Further, from $\callCond{\Phi_S,r_1,r_2,a}$ and $b \leq a$ and $a + \calllen -1 \leq e$ and $[b,e] \subseteq \ta$ we can conclude that $\Phi_S$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$.
+    Further, from $\callCond{\Phi_S,r_1,r_2,\offpc,\offsigma,a}$ and $b \leq a$ and $a + \calllen -1 \leq e$ and $[b,e] \subseteq \ta$ we can conclude that $\Phi_S$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$.
     By the \emph{Guarantee stack base address before call} we then know $\Phi_S(r_\stk) = \stkptr{\_,\stkb,\_,\_}$.

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Do a check of a first part of the FTLR call case.
Particularly, highlight the place where I think we need to use the reasonability of the ec.
Fix some minor things, point out some minor problems.

@@
   In order to prove this, we first, we prove that one of the following holds:
-  \lau{TODO: add $\neq$ failed to the cases that are not failed.}
   \begin{enumproof}
@@
   By the above observation, we know that $\Phi_S \step[\gc] \Phi_S'$ and $\Phi_T \step
-  \Phi_T'$ for some $\Phi_S'$ and $\Phi_T'$. According to
-  Lemma~\ref{lem:lro-anti-red-gen}, it suffices to show:
+  \Phi_T'$ for some $\Phi_S'$ and $\Phi_T'$.
+  According to Lemma~\ref{lem:lro-anti-red-gen}, it suffices to show:
   \[
@@
 
-  In case \ref{case:ftlr:failed}, both executions go to failed. In this case, we
-  either assume that the target configuration terminates or that the source
-  configuration terminates, but in either case it leads to a contradiction with
-  the failing configuration.
+  In case \ref{case:ftlr:failed}, both executions go to $\failed$.
+  In this case, the result follows vacuously by definition of $\lrol$ and $\lror$.
   \\\\
@@
     \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
-    \end{itemize}: Follows from the corresponding assumption of this lemma and
-    the fact that $\Phi_S \nstep[1]{\ta,\stkb} \Phi_S'$ and
-    Lemma~\ref{lem:ec-reasonable-antired},
-    \ref{lem:ec-reasonable-downwards-closed}, and \ref{lem:downwards-closed}.
+    \end{itemize}
+    This follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep[1]{\gc} \Phi_S'$ and Lemma~\ref{lem:ec-reasonable-antired}, \ref{lem:ec-reasonable-downwards-closed}, and \ref{lem:downwards-closed}.
   \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$:
@@
   \item $\npair[n-1]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$:
-    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed} and the fact that integers are always in $\lrvg{\trust}$ by definition.
+    Follows from the corresponding assumption of this lemma using Lemma~\ref{lem:downwards-closed}, the definition of $\lrr$ and the fact that integers are always in $\lrvg{\trust}$ by definition.
   \item $\memSat[n-1]{\Phi_S.\mem,\Phi_S.\ms_\stk,\Phi_S.\stk,\Phi_T.\mem}{W_M}$:
@@
   In case \ref{case:ftlr:call}, we may assume $r_1 \neq r_2$\footnote{If the register contains a data capability, then the execution fails in the step after the jump.
-    If it is an executable capability, then the xjump fails as it does not permit executable capabilities for the data part.}
-  and $r_i \neq \pcreg$ for $i \in \{1,2\}$\footnote{The pc is executable which causes the xjump to fail.}.
-  as this will cause the execution to fail.
+If it is an executable capability, then the xjump fails as it does not permit executable capabilities for the data part.} and $r_i \neq \pcreg$ for $i \in \{1,2\}$\footnote{The pc is executable which causes the xjump to fail.} as this will cause the execution to fail.
   We need to let the target execution catch up.
@@
   \]
-  By assumption, we have $\npair[n]{(\reg_S(r_i),\reg_T(r_i))} \in \lrrg{\trust}(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
+  By assumption, we have $\npair[n]{(\reg_S(r_i),\reg_T(r_i))} \in \lrvg{\trust}(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
   We know the capabilities in $r_1$ and $r_2$ are sealed capabilities, and by
@@
   \end{equation}
-  Now construct $W_{R,1}'$ as follows:
-
-  By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r$ in $S$ $\pwfree[W_{R,1}](r).H \nequal \stdreg{R(r),\gc}{\pur}.H$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
-  Now define
+  Now take $W_o = W_{R,2}$, take $n' = n-1$ and construct $W_{R,1}'$ as follows:
+
+  By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r \in S$, $\pwfree[W_{R,1}](r).H \nequal \stdreg{R(r),\gc}{\pur}.H$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
+  Now take $r_\var{priv\_stk}$ fresh and define
   \[
@@
   \]
-  For a fresh $r_\var{priv\_stk}$.
   We know $W_{R,1}' \future W_{R,1}$ as the revoked regions must have been $\spatial$ in $W_{R,q}$ (as they are owned by the part of the world assigned to the stack-register in the register-file relation).
@@
 
-  Pick this world as $W'$ in Eq~\ref{eq:ftlr:call:sealed-code-assump}
-  . Let $W_{R,2}'$ be the same world but
-  with the ownership of $W_{R,2}$ and pick it for $W_o$.
-  Now use monotonicity of $H_\sigma$ with the above facts to get
+  Pick this world as $W'$ in Eq~\ref{eq:ftlr:call:sealed-code-assump}.
+  Let $W_{R,2}'$ be the same world but with the ownership of $W_{R,2}$ and pick it for $W_o$.
+  Now observe that also $W_{R,2}' \future W_{R,2}$ and use monotonicity of $H_\sigma$ with the above facts to get
   \[
@@
   \]
+
   Now pick register files and memories such that they form $\Phi_S''$ (defined in
@@
   \begin{enumerate}[label=\roman*.]
-  \item \label{case:ftlr:scall:world} $W_{R,1}'\oplus W_{R,2}' \oplus W_R' \oplus W_M'$
+  \item \label{case:ftlr:scall:world} $W_{R,1}'\oplus W_{R,2}' \oplus W_R' \oplus W_M'$ is defined.
   \item \label{case:ftlr:scall:reg} $\npair[n-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$
@@
   as desired.
+
   It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}, but first we note that we can deduce the following: From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\sigrets',\_,M(r),\gc}$.
@@
   where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
-  That is $\mscode'$ contains the call we are considering.
+  \dominique{How do we know $\ms_\code'$ contains [b,e] entirely?
+    It does seem plausible that it contains the call entirely as well as $a + \offpc$ thanks to the code region's well-formedness.
+  }
+  That is: $\mscode'$ contains the call we are considering.
   This also entails
@@
   Case \ref{case:ftlr:scall:reg}: We need to split the ownership of $W_R'$.
-  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
+  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we get a way to split the ownership of $W_R$.
   We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership (i.e.\ $\purePart{W_{R,1}'}$).
@@
   result follows from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity}.
+  \dominique{No: we need to use Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} together with the reasonability of registers at a boundary crossing (if there is one).}
+
   This leaves the following cases:
@@
     In the next case, it is required that $\sigma' \in \sigcloss'$, but this
-    cannot be the case as $\sigma' \in \sigrets'$ and we have $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$.
-\lau{TODO I think this case should follow from $H^\var{code}$ and the comp-code judgement, but it does not.
-      Specifically, the comp-code judgement (for now) does not require $\sigrets'$ and $\sigcloss'$ to be disjoint.}
+    cannot be the case as $\sigma' \in \sigrets'$ and we have $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$, which implies that $\sigcloss' \mathrel{\#} \sigrets'$.
 

### 2018-06-25 fully-abstract-comp/technical-report.tex
Check some easy lemmas.
Make callCondition mention offpc and offsigma.
Do a check of a first part of the FTLR call case.
Particularly, highlight the place where I think we need to use the reasonability of the ec.
Fix some minor things, point out some minor problems.

@@
   \sourcecolor\left.
-    \callCond{\Phi,r_1,r_2,\aaddr} = \left\{
+    \callCond{\Phi,r_1,r_2,\offpc,\offsigma,\aaddr} = \left\{
       \begin{array}{l}
@@
 = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
-                                                    \callCond{\Phi,r_1,r_2,\aaddr} \tand\\
+                                                    \callCond{\Phi,r_1,r_2,\offpc,\offsigma,\aaddr} \tand\\
                                                     \lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta \tand \\
@@
                                                     \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \tand \\
-                                                    \src{(\neg\callCond{\Phi,r_1,r_2,\aaddr} \tor}\\
+                                                    \src{(\neg\callCond{\Phi,r_1,r_2,\offpc,\offsigma,\aaddr} \tor}\\
                                                     \src{\eaddr < \aaddr + \calllen-1) \tand} \\
@@
       \quad \uplus_{r \in S} P(r) \supseteq [\baddr,\eaddr] \\
-      \quad \forall r \in S, W'\ldotp \npair{\stpair{\ms}{\ms}} \in \pwheap(r).H \; W'  \ldotp \\
+      \quad \forall r \in S, n' < n, W'\ldotp \npair{\stpair[n']{\ms}{\ms}} \in \pwheap(r).H \; W'  \ldotp \\
       \qquad \dom(\src{\ms_S}) = P(r) \wedge \dom(\ms_T) = P(r)
@@
   \label{def:reasonable-conf}
-  We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb)$ 
+  We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
   iff for all $n' \le n$:
+  \dominique{require that evaluations below stay within \ta or even within the bounds of the current pc?}
   \begin{itemize}
@@
 \[
-  \callCond{\Phi,r_1,r_2,\aaddr}
+  \callCond{\Phi,r_1,r_2,\offpc,\offsigma,\aaddr}
 \]

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Continue pass over call case of FTLR.

@@
     is indeed the case.
+
     Finally we need to show:
@@
     \]
-    To this end let $W'' \future \purePart{W_{R,\rretc}'}$ and $W_o$ be given s.t.
-    $W'' \oplus W_o$ is defined.
+
+    To this end let $W'' \future \purePart{W_{R,\rretc}'}$ and $W_o$ be given s.t. $W'' \oplus W_o$ is defined.
     Further, let $\npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o)$ be given and show
@@
     \]
+
     Now let $n''' \leq n''$ be given along with $\reg[3]_S$, $\reg[3]_T$, $\ms^{(3)}_S$, $\ms^{(3)}_T$,$\ms^{(3)}_\stk$, $\stk^{(3)}$, $W_R''$, and $W_M''$ such that
@@
     We know that the two register-files are related which in particular means that the values in register $\rstk$ are related.
-    Now consider the following cases: \lau{TODO: Perhaps make a lemma that says if source fail and target step to fail, then the two configurations are in $\lro$.}
+    Now consider the following cases: 
     \begin{itemize}
@@
       In all cases, the source configuration will fail because the value in the stack register is not a stack capability.
-      If we can argue that the target configuration also will fail, then the two are in the observation relation.
+      If we can argue that the target configuration will also fail, then the two are in the observation relation.
       First, if $\reg[3]_T = \sealed{\sigma_\var{ret\_\stk},\vsc_T''}$, then the return code will fail when the base address (a sealed capability has no base address, so the instruction returns $-1$) is compared to $\stkb$.
-      Second, if $\reg[3]_T = \seal{\_,\_,\_}$, then the target execution fails when it attempts to splice this seal with $\vsc_T'$.
+      Second, if $\reg[3]_T = \seal{\_,\_,\_}$, then the target execution fails when it attempts to splice this seal with $\vsc_T'$ (which we know is not a $\seal{}$ capability).
       Finally, $\reg[3]_T = ((\perm_\var{ret\_\stk},\lin_\var{ret\_\stk}),\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_)$ and $\npair[n''']{[\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk}]} \in \readCond{\lin_\var{ret\_\stk},W_{R,\rstk}''}$ is satisfied.
@@
       This means that the execution will fail that $\stkb$ check.
+
     \item $\reg_S^{(3)}(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\_,\_}$ and $\baddr_\var{ret\_\stk} \neq \stkb$\\
@@
       Similarly on the target side, the return code will fail the $\stkb$ check.
+
     \item $\reg_S^{(3)}(\rstk) = \stkptr{\_,\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and either $\eaddr_\var{ret\_\stk} +1 \neq \baddr_\stk' $ or $\perm_\var{ret\_\stk} \neq \rw$ or
@@
       On the target side, the splice will fail as either the two capabilities being spliced don't line up, the permissions don't match, or the range of authority is empty, respectively.
+
     \item $\reg_S^{(3)}(\rstk) = \stkptr{\perm_\var{ret\_\stk},\baddr_\var{ret\_\stk},\eaddr_\var{ret\_\stk},\_}$ and $\baddr_\var{ret\_\stk} = \stkb$ and $\eaddr_\var{ret\_\stk} +1 = \baddr_\stk' $ and $\perm_\var{ret\_\stk} = \rw$ and $\baddr_\var{ret\_\stk} \leq \eaddr_\var{ret\_\stk}$.
@@
       \]
+
       Assume for contradiction $\ms_{\priv,S}^{(3)}$ is not the top frame. In that
@@
       Therefore, the top stack frame must contain $\ms_{\priv,S}^{(3)}$.
+
       Further, due to the disjointedness required by memory satisfaction, it must be $r'$ that governs this stack frame.
@@
         \]
-        and $W'''_R$ with the ownership of $W''_R$ as well as for the regions $r_{b_\stk'}\dots r_{e_\stk'}$,
-        and $W'''_M$ with the ownership of $W''_M$,
+        and $W'''_R$ the same as $W'''$, but with the ownership of $W''_R$ as well as for the regions $r_{b_\stk'}\dots r_{e_\stk'}$,
+        and $W'''_M$ the same as $W'''$ but with the ownership of $W''_M$,
         we show the following:
         \begin{enumproof}
-        \item $\npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrv(W'''_R)$
+        \item $\npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrr(W'''_R)$
         \item $\memSat[n''']{\Phi_S^{(4)}.\mem,\Phi_S^{(4)}.\stk,\Phi_S^{(4)}.\ms_\stk,\Phi_T^{(5)}.\mem}{W'''_M}$ 
@@
         \end{enumproof}
-        Assuming the above, we use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$ (note that from assumption ``$\Phi_S$ reasonable up to $n$ steps'' and Lemma~\ref{lem:ec-reasonable-downwards-closed} and $n''' \le n'' < n - 1$ we get ``$\Phi_S$ reasonable up to $n'''+1$ steps'' from which it follows that ``$\Phi_S(\pcreg) + \calllen$ reasonable up to $n'''$ steps''.
-        ) To get
-        \[
-          \npair[n''']{(((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen))} \in \lre[\trusted](W''')
-        \]
+
+        Assuming the above, we use our assumption that Theorem~\ref{thm:ftlr} holds for all $n' < n$ to get
+        \[
+          \npair[n''']{(((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\calllen))} \in \lre(W''')
+        \]
+        Note that from assumption ``$\Phi_S$ reasonable up to $n$ steps'' and Lemma~\ref{lem:ec-reasonable-downwards-closed} and $n''' \le n'' < n - 1$ we get ``$\Phi_S$ reasonable up to $n'''+1$ steps'' from which it follows that ``$\Phi_S(\pcreg) + \calllen$ reasonable up to $n'''$ steps''.
         Using this along with the register-file safety and memory satisfaction,
@@
         \[
-          \npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrrtrusted(W'''_R)
-        \]
-        Using Lemma~\ref{lem:untrusted-subset-trust}, it suffices to show $\npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrrtrusted(W'''_R)$.
+          \npair[n''']{(\Phi_S^{(4)}.\reg,\Phi_T^{(5)}.\reg)} \in \lrr(W'''_R)
+        \]
         We will split the world like in $\npair[n''']{(\reg_S^{(3)},\reg_T^{(3)})} \in \lrr(W)$, but where $r_\stk$ also get the ownership of $r_{b_\stk'}\dots r_{e_\stk'}$.
@@
           We already know:
-          \[\npair[n''']{(\reg^{(3)}_S(r),\reg^{(3)}_T(r))} \in \lrr(W''_{R,r}),\] so by Lemma~\ref{lem:monotonicity}, we get
+          \[\npair[n''']{(\reg^{(3)}_S(r),\reg^{(3)}_T(r))} \in \lrr(W''_{R,r}),\]
+          This is true for some $W_{R,r}''$ which does not have the ownership of the regions that are revoked in $W_R'''$, so that we can take the corresponding $W_{R,r}'''$ and have $W_{R,r}''' \future W_{R,r}''$.
+          From Lemma~\ref{lem:monotonicity}, we then get
           \[
@@
 
-          The $\var{stackReadCondition}$ is showed in the same way, but we also use Lemma~\ref{lem:stdreg-singleton-addr-strat} to argue that the new regions are address-stratified.
+          The $\var{stackReadCondition}$ is shown in the same way, but we also use Lemma~\ref{lem:stdreg-singleton-addr-strat} to argue that the new regions are address-stratified.
         \end{itemize}
@@
 
-          Follows by region interpretation monotonicity and the heap satisfaction assumption.
+          Follows by Lemma~\ref{lem:monotonicity} and the heap satisfaction assumption.
         \end{itemize}
@@
 This concludes case \ref{case:ftlr:scall:reg}
-
-% Case: \ref{case:ftlr:scall:mem}, we need to show
-
-% \begin{equation}
-%   \memSat[n-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}\label{eq:ftlr:call:memsat-case}
-% \end{equation}
-% We have the following memory satisfaction on the initial memory:
-% \begin{equation}
-%   \memSat[n]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}\label{eq:ftlr-call-og-mem-sat}
-% \end{equation}
-% By definition of memory satisfaction, this gives us:
-% \begin{itemize}
-% \item $\stk = (\opc_1,\ms_1) :: \dots :: (\opc_m,\ms_m)$
-% \item $\ms_1 \# \dots \# \ms_m \# \ms_S \# \ms_\stk$
-% \item $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$
-% \item $\ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}$ such that
-%   \begin{itemize}
-%   \item $\ms_S = \ms_{f,S} \uplus \ms_S'$
-%   \item $\ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
-%   \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}}$
-%   \item $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$
-%   \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})$
-%   \end{itemize}
-% \end{itemize}
-% We need to consider the memories of $\Phi_S''$ and $\Phi_T''$.
-% The memories of the initial configurations have only changed little:
-% \[
-%   \Phi_T''.\mem = \Phi_T''.\mem\update{\aaddr_\stk}{42}
-% \]
-% and 
-% \[
-% \Phi_S''.\ms_\stk = \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
-% \]
-% and
-% \[
-%   \Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk
-% \]
-% with $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}$.
-% We need to show the memory satisfaction in \ref{eq:ftlr:call:memsat-case}. Say
-% $(\opc_0,\ms_0) = ((\aaddr+\calllen),\ms_{\stk\_\priv,S})$. In this case we have
-% to argue 
-% \[
-%   \ms_{\stk\_\priv,S} \# \ms_1 \# \dots \# \ms_m \# \ms_S \# \Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]}
-% \]
-% which follows from Assumption~\ref{eq:ftlr:call:memsat-case} and the fact that the memory we remove from $\Phi_S.\ms_\stk$ is exactly what we add to the private stack.
-% \lau{This is a bit more involved than I make it to be here, but the fact that we also have a safety assumption on the register-file with the stack pointer means that all of the memory removed from the free stack is actually there.}
-
-%   This concludes case \ref{case:ftlr:scall:reg}
 
@@
   \item $\memSatFStack{\Phi_S.\ms_\stk,\ms_\var{T,free\_stack}}{W_{M,\var{free\_stack}}}$
-  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{M,\var{heap}})$
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap[W_M])(W_{M,\var{heap}})$
   \end{itemize}
@@
 
-    By assumption we know $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$ which entails $\npair{(\Phi_S.\reg(\rstk),\Phi_T.\reg(\rstk))} \in \lrrg{\trust}(W_{R,\rstk})$.
+    By assumption we know $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrr(W_R)$ which entails $\npair{(\Phi_S.\reg(\rstk),\Phi_T.\reg(\rstk))} \in \lrr(W_{R,\rstk})$.
     We know $\Phi_S.\reg(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk}$, so by the definition of $\lrv$ we get
@@
     We know that all regions in $S_{\rstk}$ govern singleton memory segments, so $R_\ms$ must map to singleton memory segment pairs for $r \in S_{\rstk}$.
-    Further, by definition of the free stack satisfaction for $r \in \S_{\rstk}$ we have
+    Further, by definition of the free stack satisfaction for $r \in S_{\rstk}$ we have
     \[
@@
     \[
-      \npair[n-1]{R_\ms(r)} \in \stdreg{R_{\rstk}(r),\gc}{\spao} \; \xi^{-1}(R_W(r))
+      \npair[n-1]{R_\ms(r)} \in \stdreg{R_{\rstk}(r),\gc}{\spao}.H \; \xi^{-1}(R_W(r))
     \]
@@
 
-  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{M,\var{heap}}')$:
+  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap[W_M'])(W_{M,\var{heap}}')$:
 
@@
   \label{lem:heap-sat-unchanged-future}
-  If $\pwheap = \pwheap[W']$ and $W' \future W$ and $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap(W)$, then $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in\lrheap(W')$.
+  If $\pwheap[\purePart{W}] = \pwheap[\purePart{W'}]$ and $W_{h}' \future W_h$ and $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap(\pwheap[W])(W_h)$, then $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in\lrheap(\pwheap[W'])(W_h')$.
 \end{lemma}

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
compilation issue fixed

@@
   \item Case \ref{case:ftlr:move-cca-etc:move} and $\isLinear{w_2}$ and $r_2 = \pcreg$.
-  \item Case \ref{case:ftlr:move-cca-etc:cseal} and $\isLinear{\Phi_S(r_1)}$ and $r_1 = \pcreg$.
+  \item Case \ref{case:ftlr:move-cca-etc:cseal} and $\isLinear{\Phi_S(r_1)}$ and
+    $r_1 = \pcreg$.
+  \item Case \ref{case:ftlr:move-cca-etc:split-normal} and
+    $\isLinear{\Phi_S(r_3)}$ and $\pcreg = r_3$.
   \end{itemize}
@@
   \item Case \ref{case:ftlr:move-cca-etc:split-normal}:
+    First we only consider registers $r_1,r_2,r_3$.
+
+    If $\isLinear{(\Phi_S(r_3))}$, then we need to redistribute the ownership
+    previously owned by the world used for $r_3$ to $r_1$ and $r_2$.
+
+    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, we get that either $\perm = \noperm$, in which case we can pick the ownership distribution to be the same as before, or $\npair{[\baddr,\eaddr]} \in \readCond{\linear,R_R(r_3)}$.
+
+    In the latter case, we get $S \subseteq \addressable{\linear,R_R(r_3)}$ and $R : S \fun \powerset{R_R(r_3)}$ such that $\forall r \in S \ldotp |R(r)| = 1$ and $\biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr]$.
+
+    This gives us a way to distribute the ownership, namely $r_3$ gives up ownership of regions $R^{-1}([\baddr,\eaddr])$, $r_1$ gets ownership of $R^{-1}([\baddr,n])$, and $r_2$ gets ownership over $R^{-1}([n+1,\eaddr])$.
+    The remainder of the ownership distribution remains unchanged.
+
+TODO TODO TODO     
+    Now we have to argue the following:
+    \begin{itemize}
+    \item if $\perm \in \readAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,n]} \in \readCond{\linear,R_R'(r_1)}$
+      \item $\npair{[n+1,\eaddr]} \in \readCond{R_R'(r_2)}$
+      \end{itemize}
+
+    \item if $\perm \in \writeAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,n]} \in \writeCond{\linear,R_R'(r_1)}$
+      \item $\npair{[n+1,\eaddr]} \in \writeCond{\linear,R_R'(r_2)}$
+      \end{itemize}
+    \item $\perm \neq \rx$
+
+      Assuming $\perm = \rx$, then it must have been the case that $\lin = \normal$ contradicting the assumption $\isLinear{(\Phi_S(r_3))}$.
+
+    \end{itemize}
+    or
+    \begin{itemize}
+    \item 
+    \end{itemize}
+    depending on whether the permission of the capability was $\perm = \noperm$
+    or not. In both cases, the result follows immediately by the assumptions.
+
+    If $\neg \isLinear{(\Phi_S(r_3))}$, then we pick $R_R' = R_R$
 
     TODO
+
+    For regions $r \neq r_1,r_2,r_3$, we have
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R(r_1))
+    \]
+    by assumption and Lemma~\ref{lem:downwards-closed}.
   \item Case \ref{case:ftlr:move-cca-etc:split-seal}:

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Remove the lrp relation.

--- 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \begin{itemize}
-  \item If $\npair{\lin,\baddr,\eaddr} \in \lrp(W)$, then $\npair[n']{\lin,\baddr,\eaddr} \in \lrp(W)$.
   \item If $\npair{A} \in \readCond{\lin,W}$, then $\npair[n']{A} \in \readCond{\lin,W}$.
@@
 \begin{proof}
-  The properties for $\lrp$, $\readCond{}$, $\stackReadCond{}$, $\writeCond{}$, $\stackWriteCond{}$, $\execCond{}$, $\xReadCond{}$ follow easily by definition.
+  The properties for $\readCond{}$, $\stackReadCond{}$, $\writeCond{}$, $\stackWriteCond{}$, $\execCond{}$, $\xReadCond{}$ follow easily by definition.
   The property for $\lrv$ follows from the others and by definition.
@@
 \begin{proof}
-  Direct from the definition of $\sqsubseteq$ and $\lrvg{\trust}$, observing that $\npair{[b,e]} \in \readCond{\lin,W}$ implies $\npair{(\lin,b,e)}\in\lrp(W)$ by definition.
+  Direct from the definition of $\sqsubseteq$ and $\lrvg{\trust}$.
 \end{proof}
@@
 \begin{proof}
-  \dominique{Doesn't hold: lrp doesn't make sense for stack regions!}
   TODO

### 2018-06-25 fully-abstract-comp/technical-report.tex
compilation issue fixed
Remove the lrp relation.

@@
 
-\lau{RESOLVED make this only take a pair (the other uses are gone)}
-\[
-  \lrp(W) = \left\{ \npair{(\lin,\baddr,\eaddr)} \middle|
-    \begin{array}{l}
-      \exists S \subseteq \addressable{l,W} \ldotp\\
-      \exists P : S \rightarrow \powerset{\Addr} \ldotp\\
-      \quad \uplus_{r \in S} P(r) \supseteq [\baddr,\eaddr] \\
-      \quad \forall r \in S, n' < n, W'\ldotp \npair{\stpair[n']{\ms}{\ms}} \in \pwheap(r).H \; W'  \ldotp \\
-      \qquad \dom(\src{\ms_S}) = P(r) \wedge \dom(\ms_T) = P(r)
-    \end{array}
-  \right\}
-\]
+% \lau{RESOLVED make this only take a pair (the other uses are gone)}
+% \[
+%   \lrp(W) = \left\{ \npair{(\lin,\baddr,\eaddr)} \middle|
+%     \begin{array}{l}
+%       \exists S \subseteq \addressable{l,W} \ldotp\\
+%       \exists P : S \rightarrow \powerset{\Addr} \ldotp\\
+%       \quad \uplus_{r \in S} P(r) \supseteq [\baddr,\eaddr] \\
+%       \quad \lin = \linear \Rightarrow \forall r \in \ldotp S |P(r)| \wedge \\
+%       \quad \forall r \in S, n' < n, W'\ldotp \npair[n']{\stpair{\ms}{\ms}} \in \pwheap(r).H \; W'  \ldotp \\
+%       \qquad \dom(\src{\ms_S}) = P(r) \wedge \dom(\ms_T) = P(r)
+%     \end{array}
+%   \right\}
+% \]
 
@@
       \begin{array}{r l l}
-        \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
+        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
         \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
@@
       \begin{array}{r l l }
-        \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
+        % \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
         \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
@@
   iff for all $n' \le n$:
-  \dominique{require that evaluations below stay within \ta or even within the bounds of the current pc?}
+  \dominique{require that evaluations below stay within $\ta$ or even within the bounds of the current pc?}
   \begin{itemize}
@@
 
-    If $\perm = \noperm$, then $\npair{(w,w_T)} \in \lrv(W_w)$ follows if we can prove that $\npair{(\normal,b,e)} \in \lrp(W)$.
-    This follows easily by definition of $\lrp$, $\xReadCond{}$ and $\codereg{}$.
+    If $\perm = \noperm$, then $\npair{(w,w_T)} \in \lrv(W_w)$ follows directly.
 

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Prove a lemma that is simple after removing lrp.
Initial work on non-expansiveness proof.

@@
 
+\begin{lemma}[Properties of n-equality of worlds]
+  \label{lem:n-equality-props}
+  If $W_1 \nequal W_2$ then
+  \begin{itemize}
+  \item $\purePart{W_1} \nequal \purePart{W_2}$
+  \item If $W_1' \future W_1$ then there exists a $W_2'$ such that $W_2' \nequal W_1'$ and $W_2' \future W_2$.
+  \item If $W_1' \oplus W_1$ is defined, then there exists a $W_2' \nequal W_1'$ and $W_2' \oplus W_2$ is defined.
+  \item If $W_1 = W_1' \oplus W_1''$ then there exist $W_2' \nequal W_1'$ and $W_2'' \nequal W_1''$ such that $W_2 = W_2' \oplus W_2''$.
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Easy to prove by unfolding definitions and making unsurprising choices for existentially quantified worlds.
+\end{proof}
+
 \begin{lemma}[Non-expansiveness of relations]
@@
   \begin{itemize}
-  \item $\xReadCond{W} \nequal \xReadCond{W'}$.
+  \item $\readCond{W} \nequal \readCond{W'}$.
   \item If $\stackReadCond{\lin,W} \nequal \stackReadCond{\lin,W'}$.
@@
   \item If $\stackWriteCond{\lin,W} \nequal \stackWriteCond{\lin,W'}$
+  \item If $\xReadCond{W} \nequal \xReadCond{W'}$.
+  \item $\lre(W) \nequal \lre(W')$.
+  \item $\lrexj(W) \nequal \lrexj(W')$.
   \item If $\execCond{\lin,W} \nequal \execCond{\lin,W'}$.
-  \item If $\xReadCond{W} \nequal \xReadCond{W'}$.
   \item If $\lrvg{\trust}(W) \nequal \lrvg{\trust}(W')$.
   \item If $\lrrg{\trust}(W) \nequal \lrrg{\trust}(W')$.
-  \item For all $n' \le n$, $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$ iff $\memSat[n']{(\ms_S,\stk,\ms_\stk,\ms_T)}{W'}$.
+  \item If $\memSatFStack{ms_\stk,\ms_T}{W}$, then $\memSatFStack{ms_\stk,\ms_T}{W'}$.
+  \item If $\memSatStack{ms_\stk,\ms_T}{W}$, then $\memSatStack{ms_\stk,\ms_T}{W'}$.
+  \item If $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap{\pwheap}(W)$, then $\npair{(\overline{\sigma},\ms_S,\ms_T)} \in \lrheap{\pwheap[W']}(W')$.
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$ iff $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W'}$.
   \item ...
@@
 \end{lemma}
+\begin{proof}
+  The properties for $\readCond{}$, $\stackReadCond{}$, $\writeCond{}$, $\stackWriteCond{}$, $\xReadCond{}$ follow from the fact that these are defined to use the world only for comparing regions using $\nequal$, $\nsubeq$ and $\nsupeq$.
+
+  The property for $\lre$ and $\lrexj$ follow from Lemma~\ref{lem:n-equality-props}.
+
+  The property for $\execCond{}$ follows from Lemma~\ref{lem:n-equality-props} and the property for $\lre$.
+
+  The property for $\lrv$ follows from the other properties, by definition, from Lemma~\ref{lem:n-equality-props}, non-expansiveness of regions \dominique{Is $\xi$ non-expansive/contractive/?}
+  TODO
+
+  The property of $\lrr$ follows from the one for $\lrv$.
+
+  The property for $\lrheap$, $\memSatFStack{\_,\_}{\_}$, $\memSatStack{\_,\_}{\_}$ and $(\_,\_,\_,\_) :_n W$ follows from the non-expansiveness of regions in the world. \dominique{and $\xi?$}
+  TODO
+\end{proof}
 
@@
 \begin{proof}
-  TODO
+  Follows directly from the definition.
 \end{proof}
@@
 \end{proof}
+
 %%% Local Variables:

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
The FTLR does not actually need to apply to linear executable capabilities, since components aren't allowed to ever construct such capabilities.

@@
     \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
-    \end{itemize}
-  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$
+    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \end{itemize}
+  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$
   \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$
@@
     \item $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$
-    \item $\Phi_S(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+    \item $\Phi_S(\pcreg) = ((\perm,\normal),\baddr,\eaddr,\aaddr)$
     \item $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$
@@
     \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
     \end{itemize}
     This follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep[1]{\gc} \Phi_S'$ and Lemma~\ref{lem:ec-reasonable-antired}, \ref{lem:ec-reasonable-downwards-closed}, and \ref{lem:downwards-closed}.
-  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$:
+  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$:
     Follows from the definition of $\updPcAddr{}$ and the assumptions of this case.
@@
       \updReg{\rretd}{\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}} \\
-      \updReg{\rretc}{\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\retoffset)}} \\
+      \updReg{\rretc}{\sealed{\sigma',((\perm,\normal),\baddr,\eaddr,\aaddr+\retoffset)}} \\
       \updReg{\rtmp{1}}{0}\\
@@
 
-  First show: if $\lin = \normal$, then
+  First show that
   \[
@@
     \]
+
     If $[\baddr',\eaddr'] = \emptyset$, then this is trivial, so assume $[\baddr',\eaddr'] \neq \emptyset$.
@@
   \]
-  which follows from the induction hypothesis in the case where $\Phi_S'.\reg(\pcreg) = ((\rx,\lin),\baddr,\eaddr,\_)$.
+  which follows from the induction hypothesis in the case where $\Phi_S'.\reg(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$.
   In this case, the IH is applicable because we have the following:
@@
     \item $\trust = \trusted$, $\Phi_S'$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
     \end{itemize}
@@
     \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_\pcreg}$
+    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
     \end{itemize}
@@
 
-  It remains to show the case where $\Phi_S'.\reg(\pcreg) \neq ((\rx,\lin),\baddr,\eaddr,\_)$
+  It remains to show the case where $\Phi_S'.\reg(\pcreg) \neq ((\rx,\normal),\baddr,\eaddr,\_)$
 
@@
 \]
-For $c=((\rx,\lin),\baddr,\eaddr,\aaddr)$.\\
+For $c=((\rx,\normal),\baddr,\eaddr,\aaddr)$.\\
 

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Avoid reproving the FTLR in the helper lemma.

@@
     \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \item $\trust = \untrusted$ and $[b,e] \mathrel{\#} \ta$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
     \end{itemize}
@@
     \]
-
-    If $[\baddr',\eaddr'] = \emptyset$, then this is trivial, so assume $[\baddr',\eaddr'] \neq \emptyset$.
-    Further, for $n'' \leq n'$ assume
-    \begin{itemize}
-    \item $\npair[n'']{\reg_S',\reg_T'} \in \lrr(W_R')$
-    \item $\memSat{\ms_S',\stk',\ms_\stk',\ms_T'}{W_M'}$
-    \item $W' \oplus W_R' \oplus W_M'$ is defined
-    \end{itemize}
-    and show
-    \[
-      \npair[n'']{\left(
-          \begin{array}{l}
-          (\ms_S',\reg_S'\update{\pcreg}{((\rx,\normal),\baddr',\eaddr',\aaddr')},\stk',\ms_\stk') \\
-          (\ms_T',\reg_T'\update{\pcreg}{((\rx,\normal),\baddr',\eaddr',\aaddr')})
-          \end{array}
-         \right)} \in \lro
-    \]
-    To this end, we want to use the IH which means that we need the following:
-    \begin{itemize}
-    \item $\trust = \untrusted$ and $\npair[n'']{[\baddr',\eaddr']} \in \readCond{W'}$
-
-      This follows from the readCondition assumption, Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:downwards-closed}, Lemma~\ref{lem:monotonicity}, Lemma~\ref{lem:purePart-oplus}, and Lemma~\ref{lem:non-linear-pure}.
-    \item $\npair[n'']{[\baddr',\eaddr']} \in \xReadCond{W'}$
-
-      This follows from what we already showed about the readXCondition,
-      Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:downwards-closed}, and Lemma~\ref{lem:monotonicity}
-    \item $\npair[n'']{\reg_S',\reg_T'} \in \lrr(W_R')$
-
-      By assumption.
-    \item $\memSat{\ms_S',\stk',\ms_\stk',\ms_T'}{W_M'}$
-
-      By assumption.
-
-    \item $W' \oplus W_R' \oplus W_M'$ is defined
-
-      By assumption
-    \end{itemize}
+    This follows immediately from the FTLR: we know that $[b,e] \mathrel{\#} \ta$ since $\trust = \untrusted$ and we know that $n' < n - 1 < n$.
   \end{itemize}

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Eliminate some cases that are impossible because the pc is assumed to be non-linear.

@@
   \]
-  assuming that neither of the following is the case:
-  \begin{itemize}
-  \item Case \ref{case:ftlr:move-cca-etc:move} and $\isLinear{w_2}$ and $r_2 = \pcreg$.
-  \item Case \ref{case:ftlr:move-cca-etc:cseal} and $\isLinear{\Phi_S(r_1)}$ and
-    $r_1 = \pcreg$.
-  \item Case \ref{case:ftlr:move-cca-etc:split-normal} and
-    $\isLinear{\Phi_S(r_3)}$ and $\pcreg = r_3$.
-  \end{itemize}
+  Note that we know that $\Phi_S.\reg(\pcreg)$ is not linear (which will sometimes help to eliminate some cases).
 
@@
     \item TODO (the remaining cases)
-  \end{itemize}
-
-  It remains to show the case where $\Phi_S'.\reg(\pcreg) \neq ((\rx,\normal),\baddr,\eaddr,\_)$
-
-  We still need to show the following cases:
-  \begin{itemize}
-  \item Case \ref{case:ftlr:move-cca-etc:move} and $\isLinear{w_2}$ and$r_2 = \pcreg$.
-    TODO
   \end{itemize}

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Small fixes in FTLR case 5.

@@
         \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k = 1$
-      \item \label{case:ftlr:move-cca-etc:move}(move) $w_1 = \Phi_S(r_2)$, $w_1' = \Phi_S'(r_2)$, and $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg \isLinear{w_1} \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$) and $r_1 \neq r_2$ and $k = 2$.
-      \item \label{case:ftlr:move-cca-etc:move-eq}(move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_S'(r_1)$ and $k = 1$.
+      \item \label{case:ftlr:move-cca-etc:move}(move) $w_1 = \Phi_S(r_2)$, $w_1' = \Phi_T(r_2)$, and $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg \isLinear{w_1} \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$) and $r_1 \neq r_2$ and $k = 2$.
+      \item \label{case:ftlr:move-cca-etc:move-eq}(move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_T(r_1)$ and $k = 1$.
       \item \label{case:ftlr:move-cca-etc:cseal}(cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' =
@@
       \begin{cases}
-        R_R(r_2) \oplus R(r_1) & r = r_1 \\
+        R_R(r_2) \oplus R_R(r_1) & r = r_1 \\
         \purePart{W_R}         & r = r_2 \\
@@
     \[
-      \npair[n-1]{(\Phi_S'(r_1),\Phi_T'(r_1))} \in \lrvg{\trust}(R_R(r_2) \oplus R(r_1))
+      \npair[n-1]{(\Phi_S'(r_1),\Phi_T'(r_1))} \in \lrvg{\trust}(R_R(r_2) \oplus R_R(r_1))
     \]
@@
     \[
-      \npair[n-1]{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(R_R(r_2) \oplus R(r_1))
-    \]
-    this follows by assumption and Lemma~\ref{lem:monotonicity}.
+      \npair[n-1]{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(R_R(r_2) \oplus R_R(r_1))
+    \]
+    this follows by assumption and Lemma~\ref{lem:monotonicity} and~\ref{lem:oplus-future}.
 
@@
     which follows from Lemma~\ref{lem:monotonicity} and \ref{lem:world-fut-purePart} and what we have proven about the pc.
-    \lau{remember to change this if we do not manage to prove that the pc is safe.}
   \item Case \ref{case:ftlr:move-cca-etc:move-eq}:

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Some work on FTLR case 5.

@@
   Direct from the definition of $\sqsubseteq$ and $\lrvg{\trust}$.
+\end{proof}
+
+\begin{lemma}[Capability splitting retains safety for normal capabilities]
+  \label{lem:splitting-safety-normal}
+  If
+  \begin{itemize}
+  \item $\npair{(c,c)} \in \lrvg{\trust}(W)$
+  \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+  \item $b \le n \le e$
+  \item $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$
+  \item $c_2 = ((\perm,\lin),n,\eaddr,\aaddr)$
+  \item $c_3 = \linCons{c}$
+  \end{itemize}
+  then there exist $W_1,W_2$,$W_3$ such that
+  \begin{itemize}
+  \item $W = W_1 \oplus W_2 \oplus W_3$
+  \item $\npair{(c_1,c_1)} \in \lrvg{\trust}(W_1)$
+  \item $\npair{(c_2,c_2)} \in \lrvg{\trust}(W_2)$
+  \item $\npair{(c_3,c_3)} \in \lrvg{\trust}(W_3)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
+
+  From $\npair{(c,c)} \in \lrvg{\trust}(W)$, we get that either $\perm = \noperm$, in which case we can pick the ownership distribution to be the same as before, or $\npair{[\baddr,\eaddr]} \in \readCond{\linear,R_R(r_3)}$.
+
+    In the latter case, we get $S \subseteq \addressable{\linear,R_R(r_3)}$ and $R : S \fun \powerset{R_R(r_3)}$ such that $\forall r \in S \ldotp |R(r)| = 1$ and $\biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr]$.
+
+    This gives us a way to distribute the ownership, namely $r_3$ gives up ownership of regions $R^{-1}([\baddr,\eaddr])$, $r_1$ gets ownership of $R^{-1}([\baddr,n])$, and $r_2$ gets ownership over $R^{-1}([n+1,\eaddr])$.
+    The remainder of the ownership distribution remains unchanged.
+
+TODO TODO TODO     
+    Now we have to argue the following:
+    \begin{itemize}
+    \item if $\perm \in \readAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,n]} \in \readCond{\linear,R_R'(r_1)}$
+      \item $\npair{[n+1,\eaddr]} \in \readCond{R_R'(r_2)}$
+      \end{itemize}
+
+    \item if $\perm \in \writeAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,n]} \in \writeCond{\linear,R_R'(r_1)}$
+      \item $\npair{[n+1,\eaddr]} \in \writeCond{\linear,R_R'(r_2)}$
+      \end{itemize}
+    \item $\perm \neq \rx$
+
+      Assuming $\perm = \rx$, then it must have been the case that $\lin = \normal$ contradicting the assumption $\isLinear{(\Phi_S(r_3))}$.
+
+    \end{itemize}
+    or
+    \begin{itemize}
+    \item 
+    \end{itemize}
+    depending on whether the permission of the capability was $\perm = \noperm$
+    or not. In both cases, the result follows immediately by the assumptions.
+
+    If $\neg \isLinear{(\Phi_S(r_3))}$, then we pick $R_R' = R_R$
+
+    TODO
+
 \end{proof}
@@
         $w_1 = w_1' = ((\perm,\lin),\baddr,n,\aaddr)$, $w_2 = w_2' =
-        ((\perm,\lin),n+1,\eaddr,\aaddr)$, $k=2$.
+        ((\perm,\lin),n+1,\eaddr,\aaddr)$, $w_3 = w_3' = \linCons{\Phi_S(r_3)}$, $k=3$.
       \item \label{case:ftlr:move-cca-etc:split-seal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
@@
         \stkptr{\perm,n+1,\eaddr,\aaddr}$, $w_2' =
-        ((\perm,\linear),n+1,\eaddr,\aaddr)$, $k=2$.
+        ((\perm,\linear),n+1,\eaddr,\aaddr)$, $w_3 = w_3' = 0$, $k=3$.
       \item \label{case:ftlr:move-cca-etc:splice-stack}(splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
@@
     \[
-      \npair[n-1]{(\Phi_S(\pcreg),\Phi_T(\pcreg))} \in \lrvg{\trust}(R(r_1))
+      \npair[n-1]{(\Phi_S(\pcreg),\Phi_T(\pcreg))} \in \lrvg{\trust}(R_R(r_1))
     \]
@@
 
-    If $\isLinear{(\Phi_S(r_3))}$, then we need to redistribute the ownership
-    previously owned by the world used for $r_3$ to $r_1$ and $r_2$.
-
-    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, we get that either $\perm = \noperm$, in which case we can pick the ownership distribution to be the same as before, or $\npair{[\baddr,\eaddr]} \in \readCond{\linear,R_R(r_3)}$.
-
-    In the latter case, we get $S \subseteq \addressable{\linear,R_R(r_3)}$ and $R : S \fun \powerset{R_R(r_3)}$ such that $\forall r \in S \ldotp |R(r)| = 1$ and $\biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr]$.
-
-    This gives us a way to distribute the ownership, namely $r_3$ gives up ownership of regions $R^{-1}([\baddr,\eaddr])$, $r_1$ gets ownership of $R^{-1}([\baddr,n])$, and $r_2$ gets ownership over $R^{-1}([n+1,\eaddr])$.
-    The remainder of the ownership distribution remains unchanged.
-
-TODO TODO TODO     
-    Now we have to argue the following:
-    \begin{itemize}
-    \item if $\perm \in \readAllowed$, then
-      \begin{itemize}
-      \item $\npair{[\baddr,n]} \in \readCond{\linear,R_R'(r_1)}$
-      \item $\npair{[n+1,\eaddr]} \in \readCond{R_R'(r_2)}$
-      \end{itemize}
-
-    \item if $\perm \in \writeAllowed$, then
-      \begin{itemize}
-      \item $\npair{[\baddr,n]} \in \writeCond{\linear,R_R'(r_1)}$
-      \item $\npair{[n+1,\eaddr]} \in \writeCond{\linear,R_R'(r_2)}$
-      \end{itemize}
-    \item $\perm \neq \rx$
-
-      Assuming $\perm = \rx$, then it must have been the case that $\lin = \normal$ contradicting the assumption $\isLinear{(\Phi_S(r_3))}$.
-
-    \end{itemize}
-    or
-    \begin{itemize}
-    \item 
-    \end{itemize}
-    depending on whether the permission of the capability was $\perm = \noperm$
-    or not. In both cases, the result follows immediately by the assumptions.
-
-    If $\neg \isLinear{(\Phi_S(r_3))}$, then we pick $R_R' = R_R$
-
-    TODO
-
-    For regions $r \neq r_1,r_2,r_3$, we have
-    \[
-      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R(r_1))
-    \]
-    by assumption and Lemma~\ref{lem:downwards-closed}.
+    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splitting-safety-normal} gives us $W_1$, $W_2$ and $W_3$ such that $R_R(r_3) = W_1 \oplus W_2 \oplus W_3$ and $\npair{(w_1,w_1) \in \lrvg{\trust}(W_1)}$, $\npair{(w_2,w_2) \in \lrvg{\trust}(W_2)}$ and $\npair{(w_3,w_3) \in \lrvg{\trust}(W_3)}$.
+    We take $R_R'(r_1) = W_1$, $R_R'(r_2) = W_2$ and $R_R'(r_3) = W_3 \oplus R_R(r_1) \oplus R_R(r_2)$ and $R_R' = R_R(r)$ elsewhere.
+
+    By the above points, by assumption and using Lemma~\ref{lem:downwards-closed}, we then have for all $r$ that:
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R_R'(r))
+    \]
   \item Case \ref{case:ftlr:move-cca-etc:split-seal}:
-
-TODO
+    In this case, we can take $R_R' = R_R$ and use Lemma~\ref{lem:purePart-oplus} to give us that all $\purePart{R_R(r)}$ are equal.
+    For $r = r_1, r_2, r_3$, we then get easily by definition that
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R_R'(r))
+    \]
+    and for other registers, it follows by Lemma~\ref{lem:downwards-closed}.
+
   \item Case \ref{case:ftlr:move-cca-etc:split-stack}:
@@
 
-TODO
+    In this case, we can take $R_R' = R_R$ and use Lemma~\ref{lem:purePart-oplus} to give us that all $\purePart{R_R(r)}$ are equal.
+    For $r = r_1, r_2, r_3$, we then get easily by definition that
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R_R'(r))
+    \]
+    and for other registers, it follows by Lemma~\ref{lem:downwards-closed}.
   \end{itemize}

### 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
Small fix.
Formulate some additional lemmas that we will need.

@@
 
+\end{proof}
+
+\begin{lemma}[Capability splitting retains safety for stack capabilities]
+  \label{lem:splitting-safety-stack}
+  If
+  \begin{itemize}
+  \item $\npair{(c_S,c_T)} \in \lrvg{\trust}(W)$
+  \item $c_S = \stkptr{\perm,\baddr,\eaddr,\aaddr}$
+  \item $c_T = ((\perm,\linear),\baddr,\eaddr,\aaddr)$
+  \item $b \le n \le e$
+  \item $c_{1,S} = \stkptr{\perm,\baddr,n,\aaddr}$
+  \item $c_{1,T} = ((\perm,\linear),\baddr,n,\aaddr)$
+  \item $c_{2,S} = \stkptr{\perm,n,\eaddr,\aaddr}$
+  \item $c_{2,T} = ((\perm,\linear),n,\eaddr,\aaddr)$
+  \item $c_3 = 0$
+  \end{itemize}
+  then there exist $W_1,W_2$,$W_3$ such that
+  \begin{itemize}
+  \item $W = W_1 \oplus W_2 \oplus W_3$
+  \item $\npair{(c_{1,S},c_{1,T})} \in \lrvg{\trust}(W_1)$
+  \item $\npair{(c_{2,S},c_{2,T})} \in \lrvg{\trust}(W_2)$
+  \item $\npair{(c_3,c_3)} \in \lrvg{\trust}(W_3)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Similar to the proof of Lemma~\ref{lem:splitting-safety-normal}.
+\end{proof}
+
+\begin{lemma}[Capability splicing retains safety for normal capabilities]
+  \label{lem:splitting-safety-normal}
+  If
+  \begin{itemize}
+  \item $\npair{(c_1,c_1)} \in \lrvg{\trust}(W_1)$
+  \item $\npair{(c_2,c_2)} \in \lrvg{\trust}(W_2)$
+  \item $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$
+  \item $c_2 = ((\perm,\lin),n,\eaddr,\aaddr)$
+  \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+  \item $b \le n \le e$
+  \item $c_1' = \linCons{c_1}$
+  \item $c_2' = \linCons{c_2}$
+  \end{itemize}
+  then there exist $W_1',W_2',W_3'$ such that
+  \begin{itemize}
+  \item $W_1 \oplus W_2 = W_1' \oplus W_2' \oplus W_3'$
+  \item $\npair{(c,c)} \in \lrvg{\trust}(W_3')$
+  \item $\npair{(c_1',c_1')} \in \lrvg{\trust}(W_1')$
+  \item $\npair{(c_2',c_2')} \in \lrvg{\trust}(W_2')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}[Capability splicing retains safety for stack capabilities]
+  \label{lem:splitting-safety-stack}
+  If
+  \begin{itemize}
+  \item $\npair{(c_{1,S},c_{1,T})} \in \lrvg{\trust}(W_1)$
+  \item $\npair{(c_{2,S},c_{2,T})} \in \lrvg{\trust}(W_2)$
+  \item $b \le n \le e$
+  \item $c_{1,S} = \stkptr{\perm,\baddr,n,\aaddr}$
+  \item $c_{1,T} = ((\perm,\linear),\baddr,n,\aaddr)$
+  \item $c_{2,S} = \stkptr{\perm,n,\eaddr,\aaddr}$
+  \item $c_{2,T} = ((\perm,\linear),n,\eaddr,\aaddr)$
+  \item $c_S = \stkptr{\perm,\baddr,\eaddr,\aaddr}$
+  \item $c_T = ((\perm,\linear),\baddr,\eaddr,\aaddr)$
+  \item $c_{1,S}' = c_{1,T}' = 0$
+  \item $c_{2,S}' = c_{2,T}' = 0$
+  \end{itemize}
+  then there exist $W_1',W_2',W_3'$ such that
+  \begin{itemize}
+  \item $W_1 \oplus W_2 = W_1' \oplus W_2' \oplus W_3'$
+  \item $\npair{(c_S,c_T)} \in \lrvg{\trust}(W_3')$
+  \item $\npair{(c_{1,S}',c_{1,T}')} \in \lrvg{\trust}(W_1')$
+  \item $\npair{(c_{2,S}',c_{2,T}')} \in \lrvg{\trust}(W_2')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
 \end{proof}
@@
         \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
-        $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and $\lin =
-        \linear \Rightarrow w_2 = w_2' = w_3 = w_3' = 0$, and $\lin = \normal
-        \Rightarrow (w_2 = w_2' = \Phi_T(r_2) \text{ and } w_3 = w_3' =
-        \Phi_T(r_3))$, and $k=3$
+        $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and $w_2 = w_2' = \linCons{\Phi_S(r_2)}$ and $w_3 = w_3' = \linCons{\Phi_S(r_3)}$, and $k=3$
       \item \label{case:ftlr:move-cca-etc:splice-seal} (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
@@
   \item Case \ref{case:ftlr:move-cca-etc:split-normal}:
-    First we only consider registers $r_1,r_2,r_3$.
-
     From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splitting-safety-normal} gives us $W_1$, $W_2$ and $W_3$ such that $R_R(r_3) = W_1 \oplus W_2 \oplus W_3$ and $\npair{(w_1,w_1) \in \lrvg{\trust}(W_1)}$, $\npair{(w_2,w_2) \in \lrvg{\trust}(W_2)}$ and $\npair{(w_3,w_3) \in \lrvg{\trust}(W_3)}$.
@@
 
-TODO
+    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splitting-safety-stack} gives us $W_1$, $W_2$ and $W_3$ such that $R_R(r_3) = W_1 \oplus W_2 \oplus W_3$ and $\npair{(w_1,w_1) \in \lrvg{\trust}(W_1)}$, $\npair{(w_2,w_2) \in \lrvg{\trust}(W_2)}$ and $\npair{(w_3,w_3) \in \lrvg{\trust}(W_3)}$.
+    We take $R_R'(r_1) = W_1$, $R_R'(r_2) = W_2$ and $R_R'(r_3) = W_3 \oplus R_R(r_1) \oplus R_R(r_2)$ and $R_R' = R_R(r)$ elsewhere.
+
+    By the above points, by assumption and using Lemma~\ref{lem:downwards-closed}, we then have for all $r$ that:
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R_R'(r))
+    \]
+
   \item Case \ref{case:ftlr:move-cca-etc:splice-stack}:
@@
     \item $\trust = \trusted$, $\Phi_S'$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \item $\trust = \untrusted$, $[b,e] \mathrel{\#} \ta$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
     \end{itemize}
@@
     \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
-    \item $\trust = \untrusted$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \item $\trust = \untrusted$, $[b,e] \mathrel{\#} \ta$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
     \end{itemize}

### 2018-06-26 fully-abstract-comp/ftlr-proof-cases.tex
merge

@@
 TODO
+
+Copy paste from work on ftlr proof case:
+
+    First we only consider registers $r_1,r_2,r_3$.
+
+    If $\isLinear{(\Phi_S(r_3))}$, then we need to redistribute the ownership
+    previously owned by the world used for $r_3$ to $r_1$ and $r_2$.
+
+    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, we get that either $\perm = \noperm$, in which case we can pick the ownership distribution to be the same as before, or $\npair{[\baddr,\eaddr]} \in \readCond{\linear,R_R(r_3)}$.
+
+    In the latter case, we get $S \subseteq \addressable{\linear,\pwheap{R_R(r_3)}}$ and $R : S \fun \powerset{R_R(r_3)}$ such that $\forall r \in S \ldotp |R(r)| = 1$ and $\biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr]$.
+
+    This gives us a way to distribute the ownership, namely $r_3$ gives up ownership of regions $R^{-1}([\baddr,\eaddr])$, $r_1$ gets ownership of $R^{-1}([\baddr,n])$, and $r_2$ gets ownership over $R^{-1}([n+1,\eaddr])$.
+    The remainder of the ownership distribution remains unchanged.
+
+    Now we have to argue the following:
+    \begin{itemize}
+    \item if $\perm \in \readAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,s]} \in \readCond{\linear,R_R'(r_1)}$
+      \item $\npair{[s+1,\eaddr]} \in \readCond{R_R'(r_2)}$
+      \end{itemize}
+      Both follow from the assumption $\npair{[\baddr,\eaddr]} \in \readCond{\linear,R_R(r_3)}$, and the fact that $R_R'(r_i), i=1,2$ owns the appropriate regions.
+    \item if $\perm \in \writeAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,s]} \in \writeCond{\linear,R_R'(r_1)}$
+      \item $\npair{[s+1,\eaddr]} \in \writeCond{\linear,R_R'(r_2)}$
+      \end{itemize}
+      Both follow from the assumption $\npair{[\baddr,\eaddr]} \in \writeCond{\linear,R_R(r_3)}$, and the fact that $R_R'(r_i), i=1,2$ owns the appropriate regions.
+    \item $\perm = \rwx$
+
+      By assumption.
+    \item $\perm \neq \rx$
+
+      Assuming $\perm = \rx$, then it must have been the case that $\lin = \normal$ contradicting the assumption $\isLinear{(\Phi_S(r_3))}$.
+    \end{itemize}
+
+    We also trivially have $\npair[n-1]{(0,0)} \in \lrv{R_R'(r_3)}$ which shows the content of register $r_3$ to be safe.
+
+    If $\neg \isLinear{(\Phi_S(r_3))}$, then we pick $R_R' = R_R$.
+    In this case, we need to show
+    \begin{itemize}
+    \item if $\perm \in \readAllowed$, then
+      \begin{itemize}
+      \item $\npair[n-1]{[\baddr,s]} \in \readCond{\normal,R_R'(r_1)}$
+      \item $\npair[n-1]{[s+1,\eaddr]} \in \readCond{R_R'(r_2)}$
+      \end{itemize}
+      This follows by assumption, Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:non-linear-pure}, and Lemma~\ref{lem:monotonicity}.
+    \item if $\perm \in \writeAllowed$, then
+      \begin{itemize}
+      \item $\npair[n-1]{[\baddr,s]} \in \writeCond{\normal,R_R'(r_1)}$
+      \item $\npair[n-1]{[s+1,\eaddr]} \in \writeCond{\normal,R_R'(r_2)}$
+      \end{itemize}
+      This follows by assumption, Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:non-linear-pure}, and Lemma~\ref{lem:monotonicity}.
+    \item $\perm \neq \rwx$
+
+      By assumption.
+    \item if $\perm = \rx$, then 
+      \begin{itemize}
+      \item $\npair[n-1]{[\baddr,s]} \in \xReadCond{R_R'(r_1)}$
+      \item $\npair[n-1]{[s+1,\eaddr]} \in \xReadCond{R_R'(r_2)}$
+      \item $\npair[n-1]{[\baddr,s]} \in \execCond{R_R'(r_1)}$
+      \item $\npair[n-1]{[s+1,\eaddr]} \in \execCond{R_R'(r_2)}$
+      \end{itemize}
+      All of the above follow by assumption, Lemma~\ref{lem:conds-shrinkable}, Lemma~\ref{lem:non-linear-pure}, and Lemma~\ref{lem:monotonicity}.
+    \end{itemize}
+
+    Finally, we have
+    \[
+      \npair[n-1]{\Phi_S'(r_3),\Phi_T'(r_3)}\in \lrv{R_R'(3)}
+    \]
+    by assumption and Lemma~\ref{lem:downwards-closed}.
+
+  TODO
 \end{proof}
@@
 \begin{proof}
+TODO
+
+  copy paste from FTLR proof
+
+      We need to redistribute the ownership previously owned by the world used for $r_3$ to $r_1$ and $r_2$.
+
+    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, we get that either $\perm = \noperm$, in which case we can pick the ownership distribution to be the same as before, or $\npair{[\baddr,\eaddr]} \in \stackReadCond{R_R(r_3)}$.
+
+    In the latter case, we get $S \subseteq \addressable{\linear,\pwfree{R_R(r_3)}}$ and $R : S \fun \powerset{R_R(r_3)}$ such that $\forall r \in S \ldotp |R(r)| = 1$ and $\biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr]$.
+
+    This gives us a way to distribute the ownership, namely $r_3$ gives up ownership of regions $R^{-1}([\baddr,\eaddr])$, $r_1$ gets ownership of $R^{-1}([\baddr,n])$, and $r_2$ gets ownership over $R^{-1}([n+1,\eaddr])$.
+    The remainder of the ownership distribution remains unchanged.
+
+    Now we have to argue the following:
+    \begin{itemize}
+    \item if $\perm \in \readAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,s]} \in \stackReadCond{\linear,R_R'(r_1)}$
+      \item $\npair{[s+1,\eaddr]} \in \stackReadCond{R_R'(r_2)}$
+      \end{itemize}
+      Both follow from the assumption $\npair{[\baddr,\eaddr]} \in \stackReadCond{\linear,R_R(r_3)}$, and the fact that $R_R'(r_i), i=1,2$ owns the appropriate regions.
+    \item if $\perm \in \writeAllowed$, then
+      \begin{itemize}
+      \item $\npair{[\baddr,s]} \in \stackWriteCond{\linear,R_R'(r_1)}$
+      \item $\npair{[s+1,\eaddr]} \in \stackWriteCond{\linear,R_R'(r_2)}$
+      \end{itemize}
+      Both follow from the assumption $\npair{[\baddr,\eaddr]} \in \stackWriteCond{\linear,R_R(r_3)}$, and the fact that $R_R'(r_i), i=1,2$ owns the appropriate regions.
+    \end{itemize}
+
+    We also trivially have $\npair[n-1]{(0,0)} \in \lrv{R_R'(r_3)}$ which shows the content of register $r_3$ to be safe.
+
+    For regions $r \neq r_1,r_2,r_3$, we have
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R(r))
+    \]
+    by assumption and Lemma~\ref{lem:downwards-closed}.
+
   TODO
@@
       \item \label{case:ftlr:move-cca-etc:split-normal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
-        ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$, $n < \eaddr$,
+        ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $\baddr \leq s$, $s < \eaddr$,
         $w_1 = w_1' = ((\perm,\lin),\baddr,n,\aaddr)$, $w_2 = w_2' =
@@
       \item \label{case:ftlr:move-cca-etc:split-seal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
-        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \leq n$, $n <
-        \sigma_\eaddr$, $w_1 = w_1' = \seal{\sigma_\baddr,n,\sigma}$, $w_2 =
-        w_2' = \seal{n+1,\sigma_\eaddr,\sigma}$, $k=2$.
+        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \leq s$, $s <
+        \sigma_\eaddr$, $w_1 = w_1' = \seal{\sigma_\baddr,s,\sigma}$, $w_2 =
+        w_2' = \seal{s+1,\sigma_\eaddr,\sigma}$, $k=2$.
       \item \label{case:ftlr:move-cca-etc:split-stack} (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$,
-        $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$,
-        $n < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' =
-        ((\perm,\linear),\baddr,n,\aaddr)$, $w_2 =
-        \stkptr{\perm,n+1,\eaddr,\aaddr}$, $w_2' =
-        ((\perm,\linear),n+1,\eaddr,\aaddr)$, $w_3 = w_3' = 0$, $k=3$.
+        $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq s$,
+        $s < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' =
+        ((\perm,\linear),\baddr,s,\aaddr)$, $w_2 =
+        \stkptr{\perm,s+1,\eaddr,\aaddr}$,
+
+        $w_2' = ((\perm,\linear),s+1,\eaddr,\aaddr)$, $k=2$.
       \item \label{case:ftlr:move-cca-etc:splice-stack}(splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
-        $\Phi_S(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_}$, and $\Phi_T(r_3) =
+        $\Phi_S(r_2) = \stkptr{\perm,\baddr_2,\eaddr_2,\_}$, and
+
+        $\Phi_T(r_3) =
         ((\perm,\linear),\eaddr_2+1,\eaddr_3,\aaddr_3)$, $\Phi_S(r_3) =
@@
         ((\perm,\lin),\baddr_2,\eaddr_2,\_)$, and $\Phi_T(r_3) = \Phi_S(r_3) =
-        ((\perm,\lin),\eaddr_2+1,\eaddr_3,\aaddr_3)$, and $\baddr_2 \leq
+        ((\perm,\lin),\eaddr_2+1,\eaddr_3,\aaddr_3)$, and
+        $\baddr_2 \leq
         \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
@@
   which gives us $R_R : (\RegName \setminus \{\pcreg\}) \fun World$ such that $W_R = \bigoplus_{r\in \RegName \setminus \{\pcreg\}} R_R(r)$ and for all $r$ in $\RegName \setminus \{\pcreg\}$ we have $\npair[n]{(\Phi_S.\reg(r),\Phi_T.\reg(r))} \in \lrrg{\trust}(R_R(r))$.
-  To this end, we need to consider each of cases \ref{case:ftlr:move-cca-etc:cca-normal}-\ref{case:ftlr:move-cca-etc:splice-seal}:
+  To this end, we need to consider each of cases
+  \ref{case:ftlr:move-cca-etc:cca-normal}-\ref{case:ftlr:move-cca-etc:splice-seal}:
+
+  TODO: check that all the below cases take care of the trusted case.
   \begin{itemize}
@@
   \item Case \ref{case:ftlr:move-cca-etc:move}:
+    TODO: check if this case considers the trusted case.
 
@@
     and for other registers, it follows by Lemma~\ref{lem:downwards-closed}.
-
   \item Case \ref{case:ftlr:move-cca-etc:split-stack}:
+    First we only consider registers $r_1,r_2,r_3$.
 
@@
 
+    We start by arguing the safety of $r_1$:
+
+    If the permission is $\noperm$, then it is trivial to show.
+
+    Assuming the permission is not $\noperm$, we know $\perm \in \readAllowed{}$.
+
+    If we further assume $\isLinear{\Phi_S(r_2)}$
+    \begin{itemize}
+    \item $\npair{[\baddr_2,\eaddr_2]} \in \readCond{\linear,R_R(r_2)}$
+    \item $\npair{[\eaddr_2+1,\eaddr_3]} \in \readCond{\linear,R_R(r_3)}$
+    \end{itemize}
+    From this we get $S_2 \subseteq \addressable{\linear,\pwheap{R_R(r_2)}}$ and
+    $R_2 : S_2 \fun \powerset{\nats}$ such that $\forall r \in S_2 \ldotp
+    |R_2(r)| = 1$ (and so on) as well as $S_3 \subseteq \addressable{\linear,\pwheap{R_R(r_3)}}$ and
+    $R_3 : S_3 \fun \powerset{\nats}$ such that $\forall r \in S_3 \ldotp
+    |R_3(r)| = 1$ (and so on).
+
+    In $R_R'$ give $r_1$ ownership of $R_2^{-1}([\baddr_2,\eaddr_2])$ and $R_3^{-1}([\baddr_2+1,\eaddr_3])$ and removed it from the respective regions that owned it before.
+    The remaining ownership remains unchanged.
+
+    Now we need to show
+    \[
+      \npair[n-1]{[\baddr_2,\eaddr_3]} \in \readCond{\linear,R_R(r_2)}
+    \]
+    which we do by picking $S' = S_2 \cup S_3$ and $R'(r) = R_2(r)$ for $r \in S_2$ and $R'(r) = R_3(r)$ for $r \in S_3$.
+    With these choices, the rest follow by the readCondition assumptions and downwards-closure.
+
+    If $\perm \in \writeAllowed{}$, then we need to show
+    \[
+      \npair[n-1]{[\baddr_2,\eaddr_3]} \in \writeCond{\linear,R_R(r_2)}
+    \]
+    which is done in a similar fashion\footnote{We do technically need to argue that the readCondition and the writeCondition agree on the set of regions we got from the readCondition.
+      This is done using the memory satisfaction assumption.}.
+
+    Using the same ownership distribution, it is trivial to show
+    \[
+      \npair[n-1]{(\Phi_S'(r_i),\Phi_T'(r_i))} \in \lrv(R'_R(r_i)) \text{ for $i
+        \in \{1,2\}$}
+    \]
+    as the pair is $(0,0)$.
+
+    For $\neg \isLinear{\Phi_S(r_2)}$, we use $R_R' = R_R$ as the distribution function.
+    In this case, for $r_1$ it works like before, i.e.\ the result follows from the assumptions.
+    However, in the case where the permission is $\rx$, we also have to show the following:
+    \begin{itemize}
+    \item $\npair[n-1]{[\baddr_2,\eaddr_3]} \in \xReadCond{R_R'(r_1)}$
+
+      By assumption we have
+      \begin{itemize}
+      \item $\npair{[\baddr_2,\eaddr_2]} \in \xReadCond{\linear,R_R(r_2)}$
+      \item $\npair{[\eaddr_2+1,\eaddr_3]} \in \xReadCond{\linear,R_R(r_3)}$
+      \end{itemize}
+      From which the result follows.
+    \item $\npair[n-1]{[\baddr_2,\eaddr_3]} \in \execCond{R_R'(r_1)}$
+
       TODO
+    \end{itemize}
+
+    For regions $r \neq r_1,r_2,r_3$, we have
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R(r))
+    \]
+    by assumption and Lemma~\ref{lem:downwards-closed}.
+
   \item Case \label{case:ftlr:move-cca-etc:splice-seal}:

### 2018-06-26 fully-abstract-comp/ftlr-proof-cases.tex
Some work on case 6 of the FTLR

--- 2018-06-25 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \item $\perm \in \readAllowed{}$, $\perm' \in \readAllowed{}$
-  \item $\npair{(c,c')}\in\lrv(H_\sigma,W)$
+  \item $\npair{(c,c')}\in\lrvg{\trust}(H_\sigma,W)$
   \item $W \oplus W_M$ is defined
@@
   \item $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$
-  \item $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrv(W')$
-  \end{itemize}
-\end{lemma}
-\begin{proof}
+  \item $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))} \in \lrvg{\trust}(W')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Consider first the case that $\npair{(c,c')}\in\lrv(W)$.
+
   From $\npair{(c,c')}\in\lrv(H_\sigma,W)$ with $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$, $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$, $\perm \in \readAllowed{}$ and $\perm' \in \readAllowed{}$, we get that $\baddr = \baddr'$, $\eaddr = \eaddr'$ and $\aaddr = \aaddr'$ and $\npair{(\baddr,\eaddr)} \in \readCond{\lin,W}$.
@@
   \begin{itemize}
-  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
-    From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
+  \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(W)$ that also $\isLinear{\ms_T(\aaddr')}$ and $\linCons{\ms_T(\aaddr')} = 0$.
+    From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
     From the ``Additionally, if..'' case in Lemma~\ref{lem:readcond-writecond-work} with Lemma~\ref{lem:downwards-closed}, we then get that $\memSat[n']{(\ms_S\update{\aaddr}{w_S},\stk,\ms_\stk,\ms_T\update{\aaddr'}{w_T})}{W_M'}$.
 
-  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?}
+  \item case $\neg\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = \ms_S(\aaddr)$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?}
     and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
@@
   \end{itemize}
+
+  Now consider the case that $\npair{(c,c')}\in(\lrvg{\trusted}(W) \setminus \lrv(W))$. then we have that $[b,e] \subseteq \ta$ and $\npair{[b,e]} \in \xReadCond{W}$.
+  By definition of $\xReadCond{}$, there is an $r_a$ such that $\pwheap(r) \nequal \codereg{\_,\_,\code,\gc}$, $a \in \dom(\code)$.
+  By definition of $\codereg{}$ and using the fact that $\dom(\code) \ni a \in [b,e] \subseteq \ta$, we know that $\dom(\code) \subseteq \ta$ and $\_,\_\vdash_{\mathrm{comp-code}} \code$.
+  From the fact that $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$, together with the definition of $\codereg{}$, we know that $\npair[n']{(\Phi.\mem(a),\Phi.\mem(a))} \in \lrvg{\trusted}(\purePart{W_M})$ and we have that $\purePart{W_M} = \purePart{W}$ by Lemma~\ref{lem:purePart-oplus}.
+  From the fact that $\_,\_\vdash_{\mathrm{comp-code}} \code$, we get that $\neg \isLinear{\code(a)}$.
+
+  We can then take $W_M' = W_M$, $W' = W$ and get the required results from what we have proven above and using Lemma~\ref{lem:downwards-closed}.
 \end{proof}
@@
   \end{itemize}
+\end{proof}
+
+\begin{lemma}[Store to regular capability works]
+  \label{lem:readcond-cap-works}
+  If
+  \begin{itemize}
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
+  \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+  \item $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$
+  \item $\perm \in \writeAllowed{}$, $\perm' \in \writeAllowed{}$
+  \item $\aaddr \in [\baddr,\eaddr]$
+  \item $\aaddr' \in [\baddr',\eaddr']$
+  \item $\npair{(c_1,c_1')}\in\lrvg{\trusted}(H_\sigma,W)$
+  \item $\npair{(c_2,c_2')}\in\lrv(H_\sigma,W)$
+  \item $W \oplus W_M$ is defined
+  \item $n' < n$
+  \item $c_3 = \linCons{c_2}$, $c_3' = \linCons{c_2'}$
+  \end{itemize}
+  Then $\exists W', W_M'\ldotp$
+  \begin{itemize}
+  \item $W_M = W' \oplus W_M'$
+  \item $\memSat[n']{(\ms_S\update{\aaddr}{c_2},\stk,\ms_\stk,\ms_T\update{\aaddr'}{c_2'})}{W_M'}$
+  \item $\npair[n']{(c_3,c_3')} \in \lrv(W')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
 \end{proof}
@@
         $\withinBounds{w_1}$, and\\
-        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $\isLinear{w} \Rightarrow
-        w_2 = w_2' = 0$, and $\neg \isLinear{w} \Rightarrow (w_2 = \Phi_S(r_2)
-        \text{ and } w_2' = \Phi_T(r_2))$.
+        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$,  $r_2 \neq \pcreg$.
       \item (load) $w_2 = w_2' = \Phi_T(r_2) = \Phi_S(r_2) =
@@
         $w_1 = \Phi_S.\mem(\aaddr)$, and $w_1' = \Phi_T.\mem(\aaddr)$, and \\
-        $\isLinear{w_1} \Rightarrow w = w' = 0$, and \\
-        $\neg \isLinear{w_1} \Rightarrow w = w_1 \text{ and } w' = w_1'$
+        $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm}{w_1}$, $\linConsPerm{\perm}{w_1'}$ and $r_1 \neq \pcreg$
       \end{enumproof}
@@
 
-  Case \ref{case:ftlr:store-load-mem}
+  Case \ref{case:ftlr:store-load-mem}:
+
+  By Lemma~\ref{lem:lro-anti-red-gen} it suffices to show
+  \[
+    \npair[n-1]{\Phi_S',\Phi_T'} \in \lro
+  \]
+
+  First, we show that for some $W_R'$ and $W_M'$ such that $W_R \oplus W_M = W_R' \oplus W_M'$, we have that $\npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R')$ and $\memSat[n-1]{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M'}$.
+  We have that $\Phi_S' = \updPcAddr{\Phi_S\updReg{r_1,r_2}{w_1,w_2}\update{\mem.\aaddr}{w}}$ and $\Phi_T' = \updPcAddr{\Phi_T\updReg{r_1',r_2'}{w_1',w_2'}\update{\mem.\aaddr}{w'}}$ and we distinguish the following two cases:
+  \begin{itemize}
+  \item (store) $w_1 = w_1' = \Phi_S(r_1) = \Phi_T(r_1) =
+    ((\perm,\lin),\baddr,\eaddr,\aaddr)$, and $\perm \in \writeAllowed{}$,
+    and
+    $\withinBounds{w_1}$, and
+    $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$,  $r_2 \neq \pcreg$:
 
     TODO
+
+  \item (load) $w_2 = w_2' = \Phi_T(r_2) = \Phi_S(r_2) =
+    ((\perm,\lin),\baddr,\eaddr,\aaddr)$, and $\perm \in \readAllowed{}$,
+    $\withinBounds{((\perm,\lin),\baddr,\eaddr,\aaddr)}$, and
+    $w_1 = \Phi_S.\mem(\aaddr)$, and $w_1' = \Phi_T.\mem(\aaddr)$, and
+    $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm}{w_1}$, $\linConsPerm{\perm}{w_1'}$ and $r_1 \neq \pcreg$
+
+    Follows from Lemma~\ref{lem:readcond-cap-works}, redistributing ownership in the obvious way.
+
+  \end{itemize}
+
+  By the induction hypothesis, it suffices to prove that:
+  \begin{itemize}
+  \item One of the following sets of requirements holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S'$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $[b,e] \mathrel{\#} \ta$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \end{itemize}:
+
+    Follows by the same assumption of this lemma.
+
+  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$:
+
+    Follows by definition of $\updPcAddr{}$ and the fact that stores from pc into memory and loads into pc are not allowed.
+
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+
+    Follows by the same assumption of this lemma.
+
+  \item $\npair{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R')$:
+
+    See above.
+
+  \item $\memSat{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M'}$:
+
+    See above.
+
+  \item $W_\pcreg \oplus W_R' \oplus W_M'$ is defined.
+
+  \item Theorem~\ref{thm:ftlr} holds for all $n' < n-1$:
+
+    Follows by the same assumption of this lemma.
+  \end{itemize}
 

### 2018-06-26 fully-abstract-comp/ftlr-proof-cases.tex
Prove a lemma about store...

@@
   \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
-  \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-  \item $c' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$
+  \item $c_1 = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+  \item $c_1' = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$
   \item $\perm \in \writeAllowed{}$, $\perm' \in \writeAllowed{}$
@@
   \item $\aaddr' \in [\baddr',\eaddr']$
-  \item $\npair{(c_1,c_1')}\in\lrvg{\trusted}(H_\sigma,W)$
-  \item $\npair{(c_2,c_2')}\in\lrv(H_\sigma,W)$
-  \item $W \oplus W_M$ is defined
+  \item $\npair{(c_1,c_1')}\in\lrvg{\trusted}(H_\sigma,W_1)$
+  \item $\npair{(c_2,c_2')}\in\lrv(H_\sigma,W_2)$
+  \item $W_1 \oplus W_2 \oplus W_M$ is defined
   \item $n' < n$
@@
   \end{itemize}
-  Then $\exists W', W_M'\ldotp$
-  \begin{itemize}
-  \item $W_M = W' \oplus W_M'$
+  Then $\exists W_2', W_M'\ldotp$
+  \begin{itemize}
+  \item $W_2 \oplus W_M = W_2' \oplus W_M'$
   \item $\memSat[n']{(\ms_S\update{\aaddr}{c_2},\stk,\ms_\stk,\ms_T\update{\aaddr'}{c_2'})}{W_M'}$
-  \item $\npair[n']{(c_3,c_3')} \in \lrv(W')$
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  TODO
+  \item $\npair[n']{(c_3,c_3')} \in \lrv(W_2')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  From $\npair{(c_1,c_1')}\in\lrvg{\trusted}(W_1)$ and $\perm \in \writeAllowed{}$, it follows that $c_1 = c_1'$ and $\npair{[b,e]} \in \writeCond{\lin,W_1}$.
+
+  We then get a $r$ and $A$, such that $a \in A$, $\pwheap[W_1](r).H \nsupeq \stdreg{A,\gc}{v}.H$ and $\pwheap[W_1](r)$ is address-stratified.
+
+  If we decompose the judgement that $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$, then we get some $\npair[n-1]{\ms_S|_A, \ms_T|_A} \in \stdreg{A,\gc}{v}.H~W_{M,A}$ for some $W_M = W_{M,A} \oplus W_{M,R}$.
+  If we define $W_M'$ as $W_M \oplus W_2$ and $W_2' = \purePart{W_2}$, then we can use the properties about $\pwheap[W_1](r)$ above to show that  $\memSat[n']{(\ms_S\update{\aaddr}{c_2},\stk,\ms_\stk,\ms_T\update{\aaddr'}{c_2'})}{W_M'}$.
+  The fact that $\npair[n']{(c_3,c_3')} \in \lrv(W_2')$ follows from Lemma~\ref{lem:non-linear-pure}.
 \end{proof}

### 2018-06-26 fully-abstract-comp/ftlr-proof-cases.tex
Fix a TODO.
FTLR proof case 7

@@
 \begin{lemma}[Store to regular capability works]
-  \label{lem:readcond-cap-works}
+  \label{lem:store-reg-works}
   If
@@
   If we define $W_M'$ as $W_M \oplus W_2$ and $W_2' = \purePart{W_2}$, then we can use the properties about $\pwheap[W_1](r)$ above to show that  $\memSat[n']{(\ms_S\update{\aaddr}{c_2},\stk,\ms_\stk,\ms_T\update{\aaddr'}{c_2'})}{W_M'}$.
+  The fact that $\npair[n']{(c_3,c_3')} \in \lrv(W_2')$ follows from Lemma~\ref{lem:non-linear-pure}.
+\end{proof}
+
+\begin{lemma}[Store to stack capability works]
+  \label{lem:store-stack-works}
+  If
+  \begin{itemize}
+  \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$
+  \item $c_1 = \stkptr{\perm,\baddr,\eaddr,\aaddr}$
+  \item $c_1' = ((\perm',\linear),\baddr',\eaddr',\aaddr')$
+  \item $\perm \in \writeAllowed{}$, $\perm' \in \writeAllowed{}$
+  \item $\aaddr \in [\baddr,\eaddr]$
+  \item $\aaddr' \in [\baddr',\eaddr']$
+  \item $\npair{(c_1,c_1')}\in\lrvg{\trusted}(H_\sigma,W_1)$
+  \item $\npair{(c_2,c_2')}\in\lrv(H_\sigma,W_2)$
+  \item $W_1 \oplus W_2 \oplus W_M$ is defined
+  \item $n' < n$
+  \item $c_3 = \linCons{c_2}$, $c_3' = \linCons{c_2'}$
+  \end{itemize}
+  Then $\exists W_2', W_M'\ldotp$
+  \begin{itemize}
+  \item $W_2 \oplus W_M = W_2' \oplus W_M'$
+  \item $\memSat[n']{(\ms_S,\stk,\ms_\stk\update{\aaddr}{c_2},\ms_T\update{\aaddr'}{c_2'})}{W_M'}$
+  \item $\npair[n']{(c_3,c_3')} \in \lrv(W_2')$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  From $\npair{(c_1,c_1')}\in\lrvg{\trusted}(W_1)$ and $\perm \in \writeAllowed{}$, it follows that $\perm = \perm'$, $\baddr = \baddr'$, $\eaddr = \eaddr'$, $\aaddr = \aaddr'$ and $\npair{[b,e]} \in \stackWriteCond{W_1}$.
+
+  We then get a $r$ and $A$, such that $a \in A$, $\pwfree[W_1](r).H \nsupeq \stdreg{A,\gc}{v}.H$ and $\pwfree[W_1](r)$ is address-stratified.
+
+  If we decompose the judgement that $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W_M}$, then we get some $\npair[n-1]{\ms_\stk|_A, \ms_T|_A} \in \stdreg{A,\gc}{v}.H~W_{M,A}$ for some $W_M = W_{M,A} \oplus W_{M,R}$.
+  If we define $W_M'$ as $W_M \oplus W_2$ and $W_2' = \purePart{W_2}$, then we can use the properties about $\pwfree[W_1](r)$ above to show that  $\memSat[n']{(\ms_S,\stk,\ms_\stk\update{\aaddr}{c_2},\ms_T\update{\aaddr'}{c_2'})}{W_M'}$.
   The fact that $\npair[n']{(c_3,c_3')} \in \lrv(W_2')$ follows from Lemma~\ref{lem:non-linear-pure}.
@@
         $\withinBounds{w_1}$, and\\
-        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $\isLinear{w} \Rightarrow
-        w_2 = w_2' = 0$, and $\neg \isLinear{w} \Rightarrow (w_2 = \Phi_S(r_2)
-        \text{ and } w_2' = \Phi_T(r_2))$.
+        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$, $r_2 \neq \pcreg$.
       \item (load) $ w_2' = \Phi_T(r_2) =
@@
         \dom(\Phi.\ms_\stk)$, and $\aaddr \in \dom(\Phi.\ms_\stk)$, and\\
-        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and \\
-        $\isLinear{w_1} \Rightarrow w = w' = 0$, and \\
-        $\neg \isLinear{w_1} \Rightarrow w = w_1 \text{ and } w' = w_1'$
+        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm,w_1}$ and $\linConsPerm{\perm,w_1'}$ and $r_1 \neq \pcreg$.
       \end{enumproof}
@@
 
-    TODO
+    Follows from Lemma~\ref{lem:store-reg-works}, redistributing ownership in the obvious way.
 
@@
 
-  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+
+    Follows by the same assumption of this lemma, using Lemma~\ref{lem:downwards-closed}.
+
+  \item $\npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R')$:
+
+    See above.
+
+  \item $\memSat[n-1]{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M'}$:
+
+    See above.
+
+  \item $W_\pcreg \oplus W_R' \oplus W_M'$ is defined.
+
+  \item Theorem~\ref{thm:ftlr} holds for all $n' < n-1$:
 
     Follows by the same assumption of this lemma.
-
-  \item $\npair{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R')$:
+  \end{itemize}
+
+  Case \ref{case:ftlr:store-load-stack}
+
+  By Lemma~\ref{lem:lro-anti-red-gen} it suffices to show
+  \[
+    \npair[n-1]{\Phi_S',\Phi_T'} \in \lro
+  \]
+
+  First, we show that for some $W_R'$ and $W_M'$ such that $W_R \oplus W_M = W_R' \oplus W_M'$, we have that $\npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R')$ and $\memSat[n-1]{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M'}$.
+  We have that $\Phi_S' = \updPcAddr{\Phi_S\updReg{r_1,r_2}{w_1,w_2}\update{\ms_\stk.\aaddr}{w}}$, $\Phi_T' = \updPcAddr{\Phi_T\updReg{r_1',r_2'}{w_1',w_2'}\update{\ms_\stk.\aaddr}{w'}}$ and we distinguish the following two cases:
+  \begin{itemize}
+  \item (store) $w_1 = \Phi_T(r_1) =
+        ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $w_1' = \Phi_S(r_1) =
+        \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \writeAllowed{}$, and $\withinBounds{w_1}$, and $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$:
+
+        Follows from Lemma~\ref{lem:store-stack-works}, redistributing ownership in the obvious way.
+
+  \item (load) $ w_2' = \Phi_T(r_2) =
+        ((\perm,\linear),\baddr,\eaddr,\aaddr)$, and $w_2 = \Phi_S(r_2) =
+        \stkptr{\perm,\baddr,\eaddr,\aaddr}$, and $\perm \in \readAllowed{}$,
+        $\withinBounds{((\perm,\lin),\baddr,\eaddr,\aaddr)}$, and $\aaddr \in
+        \dom(\Phi.\ms_\stk)$, and $\aaddr \in \dom(\Phi.\ms_\stk)$, and\\
+        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm,w_1}$ and $\linConsPerm{\perm,w_1'}$ and $r_1 \neq \pcreg$:
+
+    Follows from Lemma~\ref{lem:load-stack-cap-works}, redistributing ownership in the obvious way.
+
+  \end{itemize}
+
+  By the induction hypothesis, it suffices to prove that:
+  \begin{itemize}
+  \item One of the following sets of requirements holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S'$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \untrusted$ and $[b,e] \mathrel{\#} \ta$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \end{itemize}:
+
+    Follows by the same assumption of this lemma.
+
+  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$:
+
+    Follows by definition of $\updPcAddr{}$ and the fact that stores from pc into memory and loads into pc are not allowed.
+
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+
+    Follows by the same assumption of this lemma, using Lemma~\ref{lem:downwards-closed}.
+
+  \item $\npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R')$:
 
@@
 
-  \item $\memSat{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M'}$:
+  \item $\memSat[n-1]{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M'}$:
 
@@
   \end{itemize}
-
-  Case \ref{case:ftlr:store-load-stack}
-
-  TODO
 

### 2018-06-27 fully-abstract-comp/technical-report.tex
The trusted value relation is *bigger* than the untrusted one.
The FTLR does not actually need to apply to linear executable capabilities, since components aren't allowed to ever construct such capabilities.
Add padding around components to prevent splicing of code pointers.
Forgot a case where the padding made a (slight) difference.

@@
                                         \withinBounds{\Phi(r_1)} \tand \\
-                                        w_2 = \linCons{\Phi(r_2)}
+                                        w_2 = \linCons{\Phi(r_2)} \tand r_2 \neq \pcreg
                                       \end{array}
@@
                                         \aaddr \in \dom(\Phi.\ms_\stk) \tand \\ % Lau: this condition seems odd as there should never be a capability for a piece of memory that is unreachable.
-                                        w_2 = \linCons{\Phi(r_2)}
+                                        w_2 = \linCons{\Phi(r_2)} \tand r_2 \neq \pcreg
                                       \end{array}\right.\\
@@
                                       \withinBounds{\Phi(r_2)} \tand \\
-                                      w = \Phi.\mem(a) \tand \\
+                                      w = \Phi.\mem(a) \tand r_1 \neq \pcreg \tand \\
                                       w_2 = \linCons{w} \tand
@@
                                       \aaddr \in \dom(\Phi.\ms_\stk) \tand \\ % Lau: this condition seems odd as there should never be a capability for a piece of memory that is unreachable.
-                                      w = \Phi.\ms_\stk(a) \tand \\
+                                      w = \Phi.\ms_\stk(a) \tand r_1 \neq \pcreg \tand \\
                                       w_2 = \linCons{w} \tand
@@
   \inference{
-    \dom(\mscode) \mathrel{\#} \dom(\msdata)&
-    \dom(\mscode) = [\baddr,\eaddr]\\
+    \dom(\mscode) = [\baddr,\eaddr]&
+    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
+    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
@@
   }{
-    \vdash (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+    \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
   }
@@
 \begin{lemma}[Untrusted is trusted]
-  \label{lem:untrusted-subset-trust}
-  \begin{itemize}
-  \item $\lrvtrusted(W) \subseteq \lrv(W)$
-  \item $\lrrtrusted(W) \subseteq \lrr(W)$
+  \label{lem:untrusted-supset-trust}
+  \begin{itemize}
+  \item $\lrvtrusted(W) \supseteq \lrv(W)$
+  \item $\lrrtrusted(W) \supseteq \lrr(W)$
   \end{itemize}
@@
   H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
-  \left\{\npair{(\code,\code)} \middle|
+  \left\{\npair{\arraycolsep=0pt\left(\array{l}\code \uplus \mspad,\\ \code \uplus \mspad\endarray\right)} \middle|
     \begin{array}{l}
-    \dom(\code) = [\aaddr,\baddr] \wedge \\
-      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
-      \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
+    \dom(\code) = [\baddr,\eaddr] \wedge \\
+      ([\baddr - 1, \eaddr + 1] \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
+      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
+      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
       \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
       \forall a \in \dom(\code)\ldotp\\
-      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
+      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})}) \wedge
     \end{array}
@@
   \[
-    \npair{\left(\arraycolsep=0pt\array{l}((\rx,\lin),\baddr,\eaddr,\aaddr),\\
-      ((\rx,\lin),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lre(W)
+    \npair{\left(\arraycolsep=0pt\array{l}((\rx,\normal),\baddr,\eaddr,\aaddr),\\
+      ((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lre(W)
   \]
@@
   \begin{itemize}
-  \item $\emptyset\neq\dom(\mscode) \mathrel{\#} \ta$
+  \item $\dom(\mscode) \mathrel{\#} \ta$
   \end{itemize}
@@
   \begin{itemize}
-  \item $\ms_S = \ms_T = \mscode$ 
-  \item $\dom(\mscode) =[\aaddr,\baddr]$
-  \item $\dom(\mscode) \subseteq \ta \vee (\dom(\mscode)\mathrel{\#} \ta \wedge \sigrets = \emptyset)$, but we know that $\emptyset \neq\dom(\mscode) \mathrel{\#} \ta$, so we get that $\sigrets = \emptyset$.
+  \item $\ms_S = \ms_T = \mscode\uplus\mspad$
+  \item $\dom(\mscode) =[\baddr,\eaddr]$
+  \item $[\baddr-1,\eaddr+1] \subseteq \ta \vee ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset)$, but we know that $\dom(\mscode) \mathrel{\#} \ta$, so we get that $\sigrets = \emptyset$.
   \item For all $a \in \dom(\code)$, $\npair{(\code(a),\code(a))} \in \lrv(\purePart{\xi(\hat{W})})$
@@
   \begin{equation*}
-    \var{comp} = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+    \var{comp} = (\mscode\uplus\mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
   \end{equation*}
+  Note that confusingly, $\var{comp}.\mscode = \mscode\uplus\mspad$, so the assumption about $\dom(\var{comp}.\mscode)$ should be interpreted properly.
   We then know from $\vdash \var{comp}$ that 
   \begin{itemize}
-  \item $\dom(\mscode) \mathrel{\#} \dom(\msdata)$
   \item $\dom(\mscode) = [\baddr,\eaddr]$
+  \item $[\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)$
+  \item $\mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]$
   \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$
@@
     \begin{equation*}
-      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
+      \npair{(\sigrets\uplus\sigcloss,\mspad\uplus\mscode\uplus \msdata', \mspad\uplus\mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
     \end{equation*}
@@
     \begin{itemize}
-    \item $R_\ms(r_\addr) = \msdata|_{\{a\}} \times \msdata|_{\{a\}}$.
-    \item $R_\ms(r_\var{code}) = \mscode\times\mscode$
+    \item $R_\ms(r_\addr) = (\msdata|_{\{a\}}, \msdata|_{\{a\}})$.
+    \item $R_\ms(r_\var{code}) = (\mspad\uplus\mscode,\mspad\uplus\mscode)$
     \end{itemize}
@@
       \end{itemize}
-    \item $\npair{(\mscode,\mscode)} \in  \codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H~ \xi^{-1}(\purePart{W'})$:
-      We take $\trust = \trusted$ if $\dom(\code) \subseteq \ta$ and $\trust =\untrusted$ otherwise.
+    \item $\npair{(\mspad\uplus\mscode,\mspad\uplus\mscode)} \in  \codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H~ \xi^{-1}(\purePart{W'})$:
+      We take $\trust = \trusted$ if $[\baddr-1,\eaddr+1] \subseteq \ta$ and $\trust =\untrusted$ otherwise.
       By definition, we need to prove that:
@@
       \begin{itemize}
-      \item $\dom(\code) = [\aaddr,\baddr]$: by assumption.
-      \item $(\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)$: by assumption and choice of $\trust$.
+      \item $\dom(\code) = [\baddr,\eaddr]$: by assumption.
+      \item $([\baddr-1,\eaddr+1] \wedge \trust = \trusted) \vee ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)$: by assumption and choice of $\trust$.
       \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code$: by assumption.
+      \item $\mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]$:
+        by assumption
       \item $\forall a \in \dom(\code), n' < n \ldotp \npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{}(\xi(\xi^{-1}(\purePart{W'}))))$:

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
Another place where padding made a slight difference.

--- 2018-06-26 fully-abstract-comp/ftlr-proof-cases.tex
@@
   It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}, but first we note that we can deduce the following: From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\sigrets',\_,M(r),\gc}$.
-  Further by $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one $r \in S$ there is a $\mscode' \subseteq \mscode$, call this $r$ for $r_{\mscode'}$,\footnote{Follows from $H^\var{code}$ requiring $\dom(\mscode') \subseteq \ta$ and $\dom(\mscode) = \ta$ and finally that $\mscode$ is part of $\ms_S$.}
+  Further by $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one $r \in S$ there is a $\mscode'\uplus\mspad \subseteq \mscode$, call this $r$ for $r_{\mscode'}$,\footnote{Follows from $H^\var{code}$ requiring $\dom(\mscode') \subseteq \ta$ and $\dom(\mscode) = \ta$ and finally that $\mscode$ is part of $\ms_S$.}
   such that
   \begin{equation}
-    \npair{(\mscode',\mscode')} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\_,\gsigrets,\gsigcloss)\label{eq:ftlr:rretd:Hcode}
+    \npair{(\mscode'\uplus\mspad,\mscode'\uplus\mspad)} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\_,\gsigrets,\gsigcloss)\label{eq:ftlr:rretd:Hcode}
   \end{equation}
@@
 \begin{itemize}
-\item $\dom(\mscode') \subseteq \ta$
-
-  Otherwise we would have $\dom(\mscode') \# \ta$ which would contradict $\dom(\mscode') \subseteq \dom(\mscode) \subseteq\ta$
+\item $\dom(\mscode'\uplus\mspad) \subseteq \ta$
+
+  Otherwise we would have $\dom(\mscode') \# \ta$ which would contradict $\dom(\mscode'\uplus\mspad) \subseteq \dom(\mscode) \subseteq\ta$
 

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
Simplify xReadCond (thanks to the padding) and prove some results about splicing.

--- 2018-06-26 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \item $\npair{(c_2,c_2)} \in \lrvg{\trust}(W_2)$
-  \item $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$
-  \item $c_2 = ((\perm,\lin),n,\eaddr,\aaddr)$
+  \item $c_1 = ((\perm,\lin),\baddr,m,\aaddr)$
+  \item $c_2 = ((\perm,\lin),m+1,\eaddr,\aaddr)$
   \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-  \item $b \le n \le e$
+  \item $b \le m \le e$
   \item $c_1' = \linCons{c_1}$
   \item $c_2' = \linCons{c_2}$
+  \item $W_1\oplus W_2 \oplus W_M$ is defined
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
   \end{itemize}
@@
 \begin{proof}
-  TODO
+  From $\npair{(c_1,c_1)} \in \lrvg{\trust}(W_1)$ and $\npair{(c_2,c_2)} \in \lrvg{\trust}(W_2)$, it follows that either ($[b,m] \mathrel{\#} \ta$ or $[b,m] \subseteq \ta$) and also either ($[m+1,e] \mathrel{\#} \ta$ or $[m+1,e] \subseteq \ta$).
+
+  Consider first the case where either $[b,m] \subseteq \ta$ or $[m+1,e]\subseteq \ta$.
+  Then by definition of $\lrvg{\trust}$, we have that $\lin = \normal$, $\trust = \trusted$ and $\npair{[b,m]} \in \xReadCond{W_1}$ or $\npair{[m+1,e]} \in \xReadCond{W_2}$, respectively.
+  It follows by definition of $\xReadCond{}$ and $\codereg{}$ that $[\baddr-1,m+1]$ or $[m,\eaddr+1] \subseteq \ta$, respectively, so that it is impossible that $[m+1,\eaddr]\mathrel{\#} \ta$ or $[\baddr,m] \mathrel{\#} \ta$ respectively.
+  In other words, we must have that both $[b,m] \subseteq \ta$ and $[m+1,e]\subseteq \ta$.
+
+  From Lemma~\ref{lem:conds-splicable}, it follows that also $\npair{[\baddr,\eaddr]}\in\xReadCond{W_1\oplus W_2}$ and from Lemma~\ref{lem:non-linear-pure}, it follows that $\npair{[\baddr,\eaddr]}\in\xReadCond{}(\purePart{W_1\oplus W_2})$.
+  Finally, we have that $\npair{(c,c)}\in\lrvg{\trust}(\purePart{W_1\oplus W_2})$.
+  We can take $W_3' = \purePart{W_1\oplus W_2}$, $W_1' = W_1$ and $W_2' = W_2$ and the remaining proof obligations follow by assumption and by Lemma~\ref{lem:purePart-duplicable} and~\ref{lem:oplus-assoc-comm}.
+
+  Now consider the case that both $[b,m] \mathrel{\#} \ta$ and $[m+1,e]\mathrel{\#} \ta$.
+  The results now follow by definition of $\lrv$, using Lemma~\ref{lem:conds-splicable} and~\ref{lem:non-linear-pure}, taking $W_3' = W_1 \oplus W_2$ and $W_1' = \purePart{W_1}$ and $W_2' = \purePart{W_2}$.
 \end{proof}
@@
     \begin{itemize}
-    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n$ steps and $[\baddr,\eaddr] \subseteq \ta$
     \item $\trust = \untrusted$ and $[b,e] \mathrel{\#} \ta$ and $\npair{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
@@
     \begin{itemize}
-    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \dom(\mscode) = \ta$
+    \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \ta$
     \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
@@
 
-  It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}, but first we note that we can deduce the following: From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $R \subseteq \addressable{\normal, \pwheap}$ and $\exists M : S \fun \powerset{\MemSeg}$ such that $\biguplus_{r\in S} \dom(M(r)) \supseteq [\baddr,\eaddr]$ and $\forall r \in S \ldotp \pwheap(r) \nequal \codereg{\sigrets',\_,M(r),\gc}$.
-  Further by $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that for one $r \in S$ there is a $\mscode' \subseteq \mscode$, call this $r$ for $r_{\mscode'}$,\footnote{Follows from $H^\var{code}$ requiring $\dom(\mscode') \subseteq \ta$ and $\dom(\mscode) = \ta$ and finally that $\mscode$ is part of $\ms_S$.}
-  such that
+  It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}, but first we note that we can deduce the following: From the assumption $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$ we get $r \in \addressable{\normal, \pwheap}$ and $\mscode$ such that $\dom(\mscode) \supseteq [\baddr,\eaddr]$ and $\pwheap(r) \nequal \codereg{\sigrets',\_,\mscode,\gc}$.
+  Further by $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we know that
   \begin{equation}
-    \npair{(\mscode',\mscode')} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\_,\gsigrets,\gsigcloss)\label{eq:ftlr:rretd:Hcode}
+    \npair{(\mscode\uplus\mspad,\mscode\uplus\mspad)} \in H^{\var{code}} \; \sigrets' \; \sigcloss' \; \mscode \; (\ta,\_,\gsigrets,\gsigcloss)~ W_r\label{eq:ftlr:rretd:Hcode}
   \end{equation}
-  where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$.
-  \dominique{How do we know $\ms_\code'$ contains [b,e] entirely?
-    It does seem plausible that it contains the call entirely as well as $a + \offpc$ thanks to the code region's well-formedness.
-  }
-  That is: $\mscode'$ contains the call we are considering.
+  where $\sigma' \in \sigrets'$ and $\dom(\mscode) \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode)$ and $W = W_r \oplus \_$.
+  That is: $\mscode$ contains the call we are considering.
   This also entails
@@
     \label{eq:ftlr:code-comp-jud}
-    \sigrets',\sigcloss' \vdash_\var{code\_comp} \mscode'
+    \sigrets',\sigcloss' \vdash_\var{code\_comp} \mscode
   \end{equation}
@@
 \begin{itemize}
-\item $\dom(\mscode') \subseteq \ta$
-
-  Otherwise we would have $\dom(\mscode') \# \ta$ which would contradict $\dom(\mscode') \subseteq \dom(\mscode) \subseteq\ta$
+\item $\dom(\mscode\uplus\mspad) \subseteq \ta$
+
+  Otherwise we would have $\dom(\mscode\uplus\mspad) \# \ta$ which would contradict $\ta\supseteq [b,e] \subseteq\dom(\mscode\uplus\mspad)$
 
@@
 
-    Use the $r=r_{\mscode'}$ as the witness.
-    The $\xReadCond{}$ gives us $\pwheap(r) \nequal \codereg{\_,\_,\mscode',\gc}$.
+    Use the $r=r_{\mscode}$ as the witness.
+    The $\xReadCond{}$ gives us $\pwheap(r) \nequal \codereg{\_,\_,\mscode,\gc}$.
     As it is a pure region, it is also present in the future world we consider.
@@
       \]
-      First, we already know $\dom(\mscode') \subseteq \ta$ and $\sigrets' \subseteq \gsigrets$.
+      First, we already know $\dom(\mscode) \subseteq \ta$ and $\sigrets' \subseteq \gsigrets$.
       Now pick $r_\var{priv\_stk}$ as the witness.
@@
       Next, $\decInstr{\mscode'(\aaddr,\aaddr+\calllen-1)} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ follows from $\callCond{\Phi_S,r_1,r_2,\offpc,\offsigma,\aaddr}$.
-      Finally, $\ms_\code(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr}$ with $\sigma' = \sigma_\baddr + \offpc \in \sigrets'$ which follows from $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$ and the fact that the call is there.
+      Finally, $\ms_\code(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr}$ with $\sigma' = \sigma_\baddr + \offpc \in \sigrets'$ which follows from $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode$ and the fact that the call is there.
     \item $\isLinear{\sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)}}$ iff
@@
     where $W_{R,\rretc}'$ has the ownership of $W$.
-    Just as in the previous case, we know that for some region $r$ there exists $\mscode' \subseteq \mscode$ such that: $\pwheap(r) \nequal \codereg{\sigrets',\_,\mscode',\gc}$ where $\sigma' \in \sigrets'$ and $\dom(\mscode') \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode')$ and $\sigrets' \subseteq \gsigrets$.
+    Just as in the previous case, we know that for some region $r$ there exists an $\mscode$ such that: $\pwheap(r) \nequal \codereg{\sigrets',\_,\mscode,\gc}$ where $\sigma' \in \sigrets'$ and $\dom(\mscode) \supseteq [\baddr,\eaddr]$ and $\aaddr + \offpc \in \dom(\mscode)$ and $\sigrets' \subseteq \gsigrets$.
     It follows easily from the definition of $H_\sigma^\var{code}$ that
@@
       \label{eq:code-caps-in-h-sigma}
-      \npair[n'']{\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr + \calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)\endarray\right)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_{R,\rretc}')
+      \npair[n'']{\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr + \calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset)\endarray\right)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_{R,\rretc}')
     \end{equation}
@@
         \forall W'' \future \purePart{W_{R,\rretc}'}, W_o, n'' < n-1,\\
-        \quad  \npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o) \ldotp \\
+        \quad  \npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o) \ldotp \\
         \qquad \npair[n'']{\retptrc(\baddr,\eaddr,\aaddr + \calllen),\vsc_S',((\rx,\normal),\baddr,\eaddr,\aaddr + \retoffset),\vsc_T'} \in \lrexj(W''\oplus W_o))
@@
     To this end let $W'' \future \purePart{W_{R,\rretc}'}$ and $W_o$ be given s.t. $W'' \oplus W_o$ is defined.
-    Further, let $\npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode' \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o)$ be given and show
+    Further, let $\npair[n'']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets' \; \sigcloss' \; \mscode \; (\ta,\stkb) \; \sigma' \; \xi^{-1}(W_o)$ be given and show
     \[
@@
     In the next case, it is required that $\sigma' \in \sigcloss'$, but this
-    cannot be the case as $\sigma' \in \sigrets'$ and we have $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode'$, which implies that $\sigcloss' \mathrel{\#} \sigrets'$.
+    cannot be the case as $\sigma' \in \sigrets'$ and we have $\sigrets',\sigcloss' \vdash_{\mathrm{comp-code}} \mscode$, which implies that $\sigcloss' \mathrel{\#} \sigrets'$.
 

### 2018-06-27 fully-abstract-comp/technical-report.tex
Discharge a simple TODO.
Simplify xReadCond (thanks to the padding) and prove some results about splicing.

@@
 \begin{proof}
-  TODO
+  Follows easily by definition.
 \end{proof}
@@
     \begin{array}{l}
-      \exists S \subseteq \addressable{\normal, \pwheap} \ldotp \\
-      \quad \exists M : S \fun \powerset{\MemSeg} \\
-      \qquad \biguplus_{r\in S} \dom(M(r)) \supseteq A \wedge\\
-      % \qquad \biguplus_{r\in S} R(r) \# \biguplus_{r\in S} C(r) \wedge \\
-      \qquad \forall r \in S \ldotp \pwheap(r) \nequal \codereg{\_,\_,M(r),\gc}
+      \exists r \in \addressable{\normal, \pwheap} \ldotp \\
+      \qquad \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\wedge\\
+      \qquad \dom(\code) \supseteq A 
     \end{array}
@@
 \]
-\lau{RESOLVED: Make this condition require a set of adjacent regions.}
 
@@
     $\npair{A'} \in \writeCond{\lin,W}$.
-  \item If $\npair{A} \in \xReadCond{W}$  and $\emptyset \neq A' \subseteq A$, then
+  \item If $\npair{A} \in \xReadCond{W}$  and $A' \subseteq A$, then
     $\npair{A'} \in \xReadCond{W}$.
@@
   The results then follow easily from the definitions.
+\end{proof}
+
+\begin{lemma}[permission-based conditions splicable]
+  \label{lem:conds-splicable}
+  If
+  \begin{itemize}
+  \item $[\baddr,\eaddr] = [\baddr_1,\eaddr_1] \uplus [\baddr_2,\eaddr_2]$
+  \item $[\baddr,\eaddr]\mathrel{\#} \ta$
+  \item $W_1\oplus W_2$ is defined.
+  \end{itemize}
+  then
+  \begin{itemize}
+  \item If $\npair{[\baddr_1,\eaddr_1]} \in \readCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \readCond{\lin,W_2}$, then $\npair{[\baddr,\eaddr]} \in \readCond{\lin,W_1 \oplus W_2}$
+  \item If $\npair{[\baddr_1,\eaddr_1]} \in \writeCond{\lin,W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \writeCond{\lin,W_2}$, then $\npair{[\baddr,\eaddr]} \in \writeCond{\lin,W_1 \oplus W_2}$.
+  \item If $\npair{[\baddr_1,\eaddr_1]} \in \execCond{W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \execCond{W_2}$, then $\npair{[\baddr,\eaddr]} \in \execCond{W_1 \oplus W_2}$.
+  \end{itemize}
+
+  If
+  \begin{itemize}
+  \item $[\baddr,\eaddr] = [\baddr_1,\eaddr_1] \uplus [\baddr_2,\eaddr_2]$
+  \item $W_1\oplus W_2 \oplus W_M$ is defined.
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
+  \end{itemize}
+  then
+  \begin{itemize}
+  \item If $\npair{[\baddr_1,\eaddr_1]} \in \xReadCond{\lin,W_1}$ and
+    $\npair{[\baddr_2,\eaddr_2]} \in \xReadCond{\lin,W_2}$, then
+    $\npair{[\baddr,\eaddr]} \in \xReadCond{\lin,W_1 \oplus W_2}$
+  \end{itemize}
+
+  If $[\baddr,\eaddr] = [\baddr_1,\eaddr_1] \uplus [\baddr_2,\eaddr_2]$, then
+  \begin{itemize}
+  \item If $\npair{[\baddr_1,\eaddr_1]} \in \stackReadCond{W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \stackReadCond{W_2}$, then $\npair{[\baddr,\eaddr]} \in \stackReadCond{W_1 \oplus W_2}$
+  \item If $\npair{[\baddr_1,\eaddr_1]} \in \stackWriteCond{W_1}$ and $\npair{[\baddr_2,\eaddr_2]} \in \stackWriteCond{W_2}$, then $\npair{[\baddr,\eaddr]} \in \stackWriteCond{W_1 \oplus W_2}$.
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  The results for $\readCond{}$, $\writeCond{}$, $\stackReadCond{}$ and $\stackWriteCond{}$ follow by taking the union of the two sets $S$, and the union of $R$ for every $r$.
+
+  The result for $\execCond{}$ follows easily by definition and using Lemmas~\ref{lem:purePart-oplus} and~\ref{lem:oplus-future}.
+
+  For $\xReadCond{}$, we get $r_i \in \addressable{\normal,W_i}$ such that $\pwheap[W_i](r_i) \nequal \codereg{\_,\_,\code_i,\gc}$ such that $\dom(\code_i)\supseteq [\baddr_i,\eaddr_i]$.
+  From $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$, it follows that $\code_i = [\baddr_i',\eaddr_i']$ and if $r_1 \neq r_2$, then $[\baddr_1'-1,\eaddr_1+1] \mathrel{\#} [\baddr_2'-1,\eaddr_2'+1]$.
+  Because $[\baddr,\eaddr] = [\baddr_1,\eaddr_1] \uplus [\baddr_2,\eaddr_2]$ and $[\baddr_i',\eaddr_i'] = \dom(\code_i)\supseteq [\baddr_i,\eaddr_i]$, we can derive that $r_1= r_2$.
+  The result then follows by definition of $\xReadCond{}$.
 \end{proof}

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
Prove another splicing lemma.
Prove the splicing-related stuff in the FTLR case 5.

@@
 \begin{lemma}[Capability splicing retains safety for stack capabilities]
-  \label{lem:splitting-safety-stack}
+  \label{lem:splicing-safety-stack}
   If
@@
   \item $\npair{(c_{2,S},c_{2,T})} \in \lrvg{\trust}(W_2)$
-  \item $b \le n \le e$
-  \item $c_{1,S} = \stkptr{\perm,\baddr,n,\aaddr}$
-  \item $c_{1,T} = ((\perm,\linear),\baddr,n,\aaddr)$
-  \item $c_{2,S} = \stkptr{\perm,n,\eaddr,\aaddr}$
-  \item $c_{2,T} = ((\perm,\linear),n,\eaddr,\aaddr)$
+  \item $b \le m \le e$
+  \item $c_{1,S} = \stkptr{\perm,\baddr,m,\aaddr}$
+  \item $c_{1,T} = ((\perm,\linear),\baddr,m,\aaddr)$
+  \item $c_{2,S} = \stkptr{\perm,m+1,\eaddr,\aaddr}$
+  \item $c_{2,T} = ((\perm,\linear),m+1,\eaddr,\aaddr)$
   \item $c_S = \stkptr{\perm,\baddr,\eaddr,\aaddr}$
   \item $c_T = ((\perm,\linear),\baddr,\eaddr,\aaddr)$
-  \item $c_{1,S}' = c_{1,T}' = 0$
-  \item $c_{2,S}' = c_{2,T}' = 0$
-  \end{itemize}
-  then there exist $W_1',W_2',W_3'$ such that
-  \begin{itemize}
-  \item $W_1 \oplus W_2 = W_1' \oplus W_2' \oplus W_3'$
-  \item $\npair{(c_S,c_T)} \in \lrvg{\trust}(W_3')$
-  \item $\npair{(c_{1,S}',c_{1,T}')} \in \lrvg{\trust}(W_1')$
-  \item $\npair{(c_{2,S}',c_{2,T}')} \in \lrvg{\trust}(W_2')$
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  TODO
+  \end{itemize}
+  then we have that
+  \begin{itemize}
+  \item $\npair{(c_S,c_T)} \in \lrvg{\trust}(W_1 \oplus W_2)$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  This follows easily by definition of $\lrv$ and using Lemma~\ref{lem:conds-splicable}.
 \end{proof}
@@
 
-    From $\npair{(\Phi_S(r_3),\Phi_S(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splitting-safety-stack} gives us $W_1$, $W_2$ and $W_3$ such that $R_R(r_3) = W_1 \oplus W_2 \oplus W_3$ and $\npair{(w_1,w_1) \in \lrvg{\trust}(W_1)}$, $\npair{(w_2,w_2) \in \lrvg{\trust}(W_2)}$ and $\npair{(w_3,w_3) \in \lrvg{\trust}(W_3)}$.
+    From $\npair{(\Phi_S(r_3),\Phi_T(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splitting-safety-stack} gives us $W_1$, $W_2$ and $W_3$ such that $R_R(r_3) = W_1 \oplus W_2 \oplus W_3$ and $\npair{(w_1,w_1) \in \lrvg{\trust}(W_1)}$, $\npair{(w_2,w_2) \in \lrvg{\trust}(W_2)}$ and $\npair{(w_3,w_3) \in \lrvg{\trust}(W_3)}$.
     We take $R_R'(r_1) = W_1$, $R_R'(r_2) = W_2$ and $R_R'(r_3) = W_3 \oplus R_R(r_1) \oplus R_R(r_2)$ and $R_R' = R_R(r)$ elsewhere.
@@
 
-TODO
+    From $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(R_R(r_2))$ and $\npair{(\Phi_S(r_3),\Phi_T(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splicing-safety-stack} tells us that $\npair{(w_1,w_1) \in \lrvg{\trust}(R_R(r_2)\oplus R_R(r_3))}$.
+    Since $w_2 = w_2'=w_3=w_3'=0$, it's clear that 
+    $\npair{(w_2,w_2') \in \lrvg{\trust}(R_R(r_1))}$ and $\npair{(w_3,w_3') \in \lrvg{\trust}(\purePart{R_R(r_3)})}$.
+    We take $R_R'(r_1) = R_R(r_2)\oplus R_R(r_3)$, $R_R'(r_2) = R_R(r_1)$ and $R_R'(r_3) = \purePart{R_R(r_3)}$ and $R_R' = R_R(r)$ elsewhere.
+
+    By the above points, by assumption and using Lemma~\ref{lem:downwards-closed}, we then have for all $r$ that:
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R_R'(r))
+    \]
+
   \item Case \ref{case:ftlr:move-cca-etc:splice-normal}:
 
-TODO
+    From $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(R_R(r_2))$ and $\npair{(\Phi_S(r_3),\Phi_T(r_3))} \in \lrvg{\trust}(R_R(r_3))$, Lemma~\ref{lem:splicing-safety-normal} gives us $W_1',W_2',W_3'$ such that $R_R(r_2)\oplus R_R(r_3) = W_1'\oplus W_2' \oplus W_3'$ and
+    $\npair{(w_i,w_i) \in \lrvg{\trust}(W_i')}$ for $i = 1,2,3$.
+    We take $R_R'(r_1) = W_1'\oplus R_R(r_1)$ (which is defined because $R_R(r_1) \oplus (R_R(r_2)\oplus R_R(r_3))$ is defined), $R_R'(r_2) = W_2'$ and $R_R'(r_3) = W_3'$ and $R_R' = R_R(r)$ elsewhere.
+
+    By the above points, by assumption and using Lemma~\ref{lem:downwards-closed}, we then have for all $r$ that:
+    \[
+      \npair[n-1]{(\Phi_S'(r), \Phi_T'(r))} \in \lrvg{\trust}(R_R'(r))
+    \]
+
   \item Case \label{case:ftlr:move-cca-etc:splice-seal}:
@@
   \]
-  which follows from the induction hypothesis in the case where $\Phi_S'.\reg(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$.
+  which follows from the induction hypothesis.
   In this case, the IH is applicable because we have the following:

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
simplify a condition.
Remember the fact that instructions like cca, restrict, mov don't modify the pc.
Small fix.
Simplify a case in the FTLR proof by not unfolding xjumpResult.
Two latex fixes
Make execution configuration reasonability into more of a "unary simulation".
Update the definition of execution configuration reasonability to achieve the
following:
* simplify the definition and make it more into a kind of "unary simulation", by
  taking one step at a time.
* make the definition more realistic, as we can hardly expect a component to
  guarantee stuff after it jumps to other components

@@
 \begin{lemma}[Capability splicing retains safety for normal capabilities]
-  \label{lem:splitting-safety-normal}
+  \label{lem:splicing-safety-normal}
   If
@@
     \item $\Phi_T \step \Phi_T'$
+    \item $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$
     \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r}{z}} \neq \failed$
@@
     \item $\Phi_T \step \Phi_T'$
+    \item $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$
     \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r_1\cdots r_k}{w_1\cdots w_k}}\neq
@@
       \neq \failed$
+    \item $r_i \neq \pcreg$ for all $i$
     \item One of the following holds:
@@
     \item $\Phi_T \step \Phi_T'$
+    \item $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$
     \item $\Phi_S' =
@@
       \updPcAddr{\Phi_T\updReg{r_1',r_2'}{w_1',w_2'}\update{\mem.\aaddr}{w'}}$
+    \item $r_i \neq \pcreg$ for all $i$
     \item One of the following hold:
@@
         $\withinBounds{w_1}$, and\\
-        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$,  $r_2 \neq \pcreg$.
+        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$.
       \item (load) $w_2 = w_2' = \Phi_T(r_2) = \Phi_S(r_2) =
@@
         $w_1 = \Phi_S.\mem(\aaddr)$, and $w_1' = \Phi_T.\mem(\aaddr)$, and \\
-        $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm}{w_1}$, $\linConsPerm{\perm}{w_1'}$ and $r_1 \neq \pcreg$
+        $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm}{w_1}$, $\linConsPerm{\perm}{w_1'}$
       \end{enumproof}
@@
       \updPcAddr{\Phi_T\updReg{r_1',r_2'}{w_1',w_2'}\update{\ms_\stk.\aaddr}{w'}}$
+    \item $r_i \neq \pcreg$ for all $i$
     \item One of the following hold:
@@
         $\withinBounds{w_1}$, and\\
-        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$, $r_2 \neq \pcreg$.
+        $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$.
       \item (load) $ w_2' = \Phi_T(r_2) =
@@
         \dom(\Phi.\ms_\stk)$, and $\aaddr \in \dom(\Phi.\ms_\stk)$, and\\
-        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm,w_1}$ and $\linConsPerm{\perm,w_1'}$ and $r_1 \neq \pcreg$.
+        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm,w_1}$ and $\linConsPerm{\perm,w_1'}$.
       \end{enumproof}
@@
     \item $\Phi_T \step \Phi_T'$
+    \item $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$
     \item One of the following holds
@@
         $\Phi_T' = \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and
-        $\isLinear{\Phi_S(r_1)} \Rightarrow w_1 = w_1' = 0$ and $\neg
-        \isLinear{\Phi_S(r_1)} \Rightarrow (w_1 = \Phi_S(r_1) \text{ and } w_1'
-        = \Phi_T(r_1))$
+        $w_1 = \linCons{\Phi_S(r_1)}$ and $w_1' = \linCons{\Phi_T(r_1)}$
       \item(xjmp) $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) =
         \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and
-        $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\Phi_S'(r_1,r_2) =
-        \linCons{c_1},\linCons{c_2}$ and $\Phi_T'(r_1,r_2) =
-        \linCons{c_1'},\linCons{c_2'}$ and\\
+        $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and
         $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
-        $\nonExec{c_\data}$ and $\Phi_T'(\pcreg,\rdata) = c_1',c_2'$ and one of
-        the following hold:
-        \begin{enumproof}
-        \item $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$ and
-          $\nonExec{c_\data}$ and $\Phi_S'(\pcreg,\rdata) = c_1,c_2$.
-        \item $c_1 = \retptrc(\baddr,\eaddr,\aaddr)$, and $c_2 =
-          \retptrd(\aaddr_\stk,\eaddr_{\stk,\priv})$ and $\Phi(r_\stk) =
-          \stkptr{\rw,\stkb,\eaddr_\stk,\_}$ and $\Phi_S.\stk = \stk_\var{frame}
-          :: (\aaddr,\ms_{\stk,\priv})$ and $\dom(\ms_{\stk,\priv} =
-          [\eaddr_\stk+1,\eaddr_{\stk,\priv}])$ and $\Phi_S'.\ms_\stk =
-          \ms_{\stk,\priv} \uplus \Phi_S.\ms_\stk$ and $\Phi_S'.\stk = \stk$ and
-          \[
-            \begin{split}
-              \Phi_S' = \Phi_S&[\pcreg \mapsto ((\rx,\normal),\baddr,\eaddr,\aaddr)]\\
-              &[\rstk \mapsto
-              \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},e_\stk+1}] \\
-              &[\rdata,\rtmp{1},\rtmp{2},r_2 \mapsto 0, 0, 0, 0]
-            \end{split}
-          \]
-        \end{enumproof}
+        $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
+        $\Phi_T'' = \Phi_T\updReg{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}$ and
+        $\Phi_T' = \xjumpResult{c_1'}{c_2'}{\Phi_T''}$
       \end{enumproof}
@@
     \item $\trust = \trusted$, $\Phi_S$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \ta$
-    \item $\trust = \untrusted$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
-    \end{itemize}
-    This follows from the corresponding assumption of this lemma and the fact that $\Phi_S \nstep[1]{\gc} \Phi_S'$ and Lemma~\ref{lem:ec-reasonable-antired}, \ref{lem:ec-reasonable-downwards-closed}, and \ref{lem:downwards-closed}.
+    \item $\trust = \untrusted$ and $[b,e] \mathrel{\#} \ta$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \end{itemize}
+    This follows from the corresponding assumption of this lemma, the fact that $\Phi_S \step[\gc] \Phi_S'$ and $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$ and \ref{lem:downwards-closed}.
   \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$:
@@
 
-  \item Case \label{case:ftlr:move-cca-etc:splice-seal}:
+  \item Case \ref{case:ftlr:move-cca-etc:splice-seal}:
 
@@
 
-    If the former holds, then it follows from Lemma~\ref{lem:ec-reasonable-antired}.
+    If the former holds, then it follows by definition of execution configuration reasonability, using the fact that $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$.
 
@@
 
-    Follows by the same assumption of this lemma.
+    Follows by the same assumption of this lemma, in the first case using the definition of execution configuration reasonability, using the fact that $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$.
 

### 2018-06-27 fully-abstract-comp/technical-report.tex
Prevent instructions like restrict, cca etc. to operate on the pc.
Make execution configuration reasonability into more of a "unary simulation".
Update the definition of execution configuration reasonability to achieve the
following:
* simplify the definition and make it more into a kind of "unary simulation", by
  taking one step at a time.
* make the definition more realistic, as we can hardly expect a component to
  guarantee stuff after it jumps to other components.
* hopefully still be strong enough to make the jmp/jnz case of the FTLR proof go
  through: components are not allowed to do

@@
                               \begin{cases}
-                                \updPcAddr{\Phi\updReg{r}{\rn}} & \rn \in \ints \\
-                                \updPcAddr{\Phi\updReg{\rn}{w}\updReg{r}{\Phi(\rn)}} & w = \linCons{\Phi(\rn)}
+                                \updPcAddr{\Phi\updReg{r}{\rn}} & r \neq \pcreg \wedge \rn \in \ints \\
+                                \updPcAddr{\Phi\updReg{\rn}{w}\updReg{r}{\Phi(\rn)}} & r \neq \pcreg \wedge w = \linCons{\Phi(\rn)}\\
+                                \failed & \totherwise
                               \end{cases}
@@
                                       \Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
-                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)
+                                      c = ((\perm,\lin),\baddr,\eaddr,\aaddr + n) \tand\\
+                                      r \neq \pcreg
                                     \end{array}
@@
                                           \decPerm{n} \sqsubseteq \perm \tand \\
-                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr)
+                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand\\
+                                          r_1 \neq \pcreg
                                         \end{array}
@@
                                                       \quad\Phi(\rn_i) = n_i \tand\\
-                                                      \quad\text{and $n_1 < n_2$}\\        
+                                                      \quad n_1 < n_2
                                                     \end{array}\\
@@
                                                       \quad \Phi(\rn_i) = n_i \tand \\
-                                                      \quad\text{and $n_1 \not< n_2$}\\        
+                                                      \quad n_1 \not< n_2\\
                                                     \end{array}\\
@@
                                       \Phi(r_1) = ((\perm,\lin),\baddr,\eaddr,\_) \tand \\
-                                      c = ((\perm,\lin),\baddr,\eaddr,\baddr)
+                                      c = ((\perm,\lin),\baddr,\eaddr,\baddr) \tand\\
+                                      r_1 \neq \pcreg
                                     \end{array} \\
@@
                                    c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr) \tand \\
-                                   w = \linCons{\Phi(r_3)}
+                                   w = \linCons{\Phi(r_3)} \tand \\
+                                   r_1,r_2,r_3 \neq \pcreg
                                  \end{array}\\
@@
                                   w_2 = \linCons{\Phi(r_2)} \tand \\
-                                  w_3 = \linCons{\Phi(r_3)} \\
+                                  w_3 = \linCons{\Phi(r_3)}  \tand \\
+                                  r_1,r_2,r_3 \neq \pcreg\\
                                 \end{array}\\
@@
   We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
-  iff for all $n' \le n$:
-  \dominique{require that evaluations below stay within $\ta$ or even within the bounds of the current pc?}
+  iff for $n' \leq n$:
   \begin{itemize}
@@
     \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$
+    \item $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$
       for some $\src{r_1}$ and $\src{r_2}$
-    \item $i \le n'$
     \end{itemize}
@@
     \begin{itemize}
-    \item $\src{\Phi'}(\src{r_\stk}) =
+    \item $\src{\Phi}(\src{r_\stk}) =
       \src{\stkptr{\_,\stkb,\_,\_}}$
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi'(\pcreg) + \calllen$ behaves reasonably up to $n'-i - 1$ steps
+    \item $\Phi(\pcreg) + \calllen$ behaves reasonably up to $n'-1$ steps
     \end{itemize}
@@
     \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $i \le n'$
-    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
+    \item $\src{\Phi}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
     \end{itemize}
@@
     \begin{itemize}
-    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$
-      and $\sigma \in \gsigrets$
+    \item $\src{\Phi}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$
     \item $\sigma \in \gsigcloss$ and one of the following holds:
       \begin{itemize}
-      \item $\exec{\Phi'(r_2)}$ and $\Phi'(r_2)$ behaves reasonably up to $n' - i - 1$ steps.
-      \item $\nonExec{\Phi'(r_2)}$ and $\Phi'(r_2)$ is reasonable up to $n' - i - 1$ steps in memory $\Phi'.\ms$ and free stack $\Phi'.\ms_\stk$.
+      \item $\exec{\Phi(r_2)}$ and $\Phi(r_2)$ behaves reasonably up to $n' - 1$ steps.
+      \item $\nonExec{\Phi(r_2)}$ and $\Phi(r_2)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$.
       \end{itemize}
@@
     \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
-    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
-    \item $i \le n'$
-    \end{itemize}
-    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n'-i -1$ steps
+    \item $\src{\Phi}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
+    \end{itemize}
+    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ step
   \item \emph{Don't leak private stuff...} If
     \begin{itemize}
-    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
-    \item $\src{\Phi'}$ is in $\ta$
-    \item $\src{\Phi''}$ is not in $\ta$
-    \item $i \le n'$
-    \end{itemize}
-    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n'-i$ steps for
-    all $r$ (including $\pcreg$)
+    \item $\Phi \step[\ta,\stkb] \Phi'$
+    \end{itemize}
+    Then one of the following holds:
+    \begin{itemize}
+    \item All of the following hold:
+      \begin{itemize}
+      \item $\Phi'.\reg(\pcreg) =
+        ((\perm,\lin),\baddr,\eaddr,\aaddr')$ and $\Phi.\reg(\pcreg) =
+        ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+      \item $\Phi'$ is reasonable up to $n'-1$ steps
+      \end{itemize}
+      % \dominique{the following seems unnecessary: why would we allow trusted code to use a jump instruction to jump out of a trusted component? it can use call and xjmp for calls/returns/tailcalls}
+    % \item $\Phi'.\reg(r)$ is reasonable in memory $\Phi'.\mem$ and free stack $\Phi'.\ms_\stk$ up to $n'-1$ steps for all $r$ (including $\pcreg$)
+    \item
+      \begin{itemize}
+      \item $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+      \item $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
+      \end{itemize}
+    \item
+      \begin{itemize}
+      \item $\Phi$ points to $\src{\txjmp{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+      \item $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
+      \end{itemize}
+    \end{itemize}
   \end{itemize}
 \end{definition}
-
-  \begin{lemma}
-    \label{lem:ec-reasonable-antired}
-    If
-    \begin{itemize}
-    \item $\Phi$ is reasonable up to $n$ steps
-    \item $\Phi \nstep[i]{\gc} \Phi'$
-    \end{itemize}
-    Then
-    \begin{itemize}
-    \item $\Phi'$ is reasonable up to $n-i$ steps
-    \end{itemize}
-  \end{lemma}
-  \begin{proof}
-    Follows easily by definition.
-  \end{proof}
 

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
Prove the jmp/jnz case of the FTLR.

@@
   \begin{enumproof}
-  \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T
-    \step[i] \failed$ for some $i$.
+  \item \label{case:ftlr:failed} $\Phi_S \nstep[j]{\gc} \failed$ and $\Phi_T
+    \step[i] \failed$ for some $i$,$j$.
   \item \label{case:ftlr:halted} $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
@@
           \sigma_{\eaddr,3}, \sigma}$
+      \item \label{case:ftlr:jnz-z} (jnz zero case) $k = 0$
       \end{enumproof}
@@
       \begin{enumproof}
-      \item(jnz) $\Phi_S' = \updPcAddr{\Phi_S}$ and $\Phi_T' =
-        \updPcAddr{\Phi_T}$
       \item(jmp,jnz) $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and
         $\Phi_T' = \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and
+        $\Phi_S(r_1) = \Phi_T(r_1) = ((\perm_1,\lin_1),\baddr_1,\eaddr_1,\aaddr_1)$,
+        $\exec{\Phi_S(r_1)}$, $\withinBounds{\Phi_S(r_1)}$,
         $w_1 = \linCons{\Phi_S(r_1)}$ and $w_1' = \linCons{\Phi_T(r_1)}$
@@
     and for other registers, it follows by Lemma~\ref{lem:downwards-closed}.
+
+  \item Case \ref{case:ftlr:move-cca-etc:jnz-z}:
+
+    In this case, we can take $R_R' = R_R$ and for all registers, the result follows by Lemma~\ref{lem:downwards-closed}.
   \end{itemize}
@@
 
+  We have that
+  \begin{itemize}
+  \item $\Phi_S \step[\gc] \Phi_S'$
+  \item $\Phi_T \step \Phi_T'$
+  \item $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$
+  \item One of the following holds
+    \begin{enumproof}
+    \item(jmp,jnz) $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and
+      $\Phi_T' = \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and
+      $\Phi_S(r_1) = \Phi_T(r_1) = ((\perm_1,\lin_1),\baddr_1,\eaddr_1,\aaddr_1)$,
+      $\exec{\Phi_S(r_1)}$, $\withinBounds{\Phi_S(r_1)}$,
+      $w_1 = \linCons{\Phi_S(r_1)}$ and $w_1' = \linCons{\Phi_T(r_1)}$
+    \item(xjmp) $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) =
+      \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and
+      $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and
+      $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
+      $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
+      $\Phi_T'' = \Phi_T\updReg{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}$ and
+      $\Phi_T' = \xjumpResult{c_1'}{c_2'}{\Phi_T''}$
+    \end{enumproof}
+  \end{itemize}
+
+  According to Lemma~\ref{lem:lro-anti-red-gen}, it suffices to show:
+  \[
+    \npair[n-1]{(\Phi_S',\Phi_T')} \in \lro
+  \]
+
+  In the first case (jmp,jnz), the fact that $\Phi_S$ is reasonable, with $\Phi_S \step[\gc] \Phi_S'$ and $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$, gives us that $\perm_1 = \perm$, $\lin_1 = \normal$, $b_1 = b$, $e_1 = e$.
+  The fact that $\lin_1 =\normal$ implies that $w_1 =\Phi_S(r_1)$ and $w_1' = \Phi_T(r_1)$.
+
+  The induction hypothesis tells us that it suffices to prove the following:
+  \begin{itemize}
+  \item One of the following sets of requirements holds:
+    \begin{itemize}
+    \item $\trust = \trusted$, $\Phi_S'$ is reasonable up to $n-1$ steps and $[\baddr,\eaddr] \subseteq \ta$
+    \item $\trust = \untrusted$ and $[b,e] \mathrel{\#} \ta$ and $\npair[n-1]{[\baddr,\eaddr]} \in \readCond{\normal,W_\pcreg}$
+    \end{itemize}
+    This follows from the corresponding assumption of this lemma, the fact that $\Phi_S \step[\gc] \Phi_S'$ and $\Phi_S$ does not point to $\scall{\offpc,\offsigma}{r_1}{r_2}$ or $\txjmp{r_1}{r_2}$ and \ref{lem:downwards-closed}.
+  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$:
+    See above.
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+    Follows by Lemma~\ref{lem:downwards-closed} from the fact that $\npair{[b,e]} \in \xReadCond{W_\pcreg}$.
+  \item $\npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R)$:
+
+    We have that $\Phi_S'.\reg(r) = \Phi_S.\reg(r)$ for all $r \neq \pcreg$, so the result follows by Lemma~\ref{lem:downwards-closed} from the corresponding assumption of this lemma.
+
+  \item $\memSat[n-1]{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M}$:
+
+    These components of $\Phi_S'$ and $\Phi_T'$ are all unmodified from $\Phi_S$ and $\Phi_T$, so the result follows by Lemma~\ref{lem:downwards-closed} from the corresponding assumption of this lemma.
+
+  \item $W_\pcreg \oplus W_R \oplus W_M$ is defined: by assumption.
+
+  \item Theorem~\ref{thm:ftlr} holds for all $n' < n-1$:
+    Follows by the same assumption of this lemma.
+  \end{itemize}
+
+  In the second case (xjmp), blabla
   TODO

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
Initial work on the FTLR xjmp case.

@@
   The fact that $\npair[n']{(c_3,c_3')} \in \lrv(W_2')$ follows from Lemma~\ref{lem:non-linear-pure}.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:sealed-lrv-lrexj}
+  If
+  \begin{itemize}
+  \item $\npair{(\sealed{\sigma,\vsc_1},\sealed{\sigma,\vsc_1'})} \in \lrvg{\trust}(W_{R,1})$
+  \item $\npair{(\sealed{\sigma,\vsc_2},\sealed{\sigma,\vsc_2'})} \in \lrvg{\trust}(W_{R,2})$ 
+  \item $\nonExec{\vsc_1'}$ and $\nonExec{\vsc_2'}$
+  \item $W_{R,1} \oplus W_{R,2}$ is defined
+  \end{itemize}
+  Then
+  \begin{itemize}
+  \item $\npair{(\vsc_1,\vsc_2,\vsc_1',\vsc_2')} \in \lrexj(W_{R,1}\oplus W_{R,2})$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  TODO
 \end{proof}
@@
         $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
+        $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$ and
         $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
@@
       $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
+      $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$ and
       $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
@@
 
-  In the second case (xjmp), blabla
+  In the second case (xjmp), we have that $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trust}(W_{R,1})$, $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(W_{R,2})$ and $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(W_R')$ for some $W_{R,1}, W_{R,2}, W_R'$ with $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$.
+
+  Using the facts that $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$, and the above points, Lemma~\ref{lem:sealed-lrv-lrexj} tells us that $\npair{(c_1,c_2,c_1',c_2')} \in \lrexj(W_{R,1} \oplus W_{R,2})$.
+
   TODO

### 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex

@@
   \item Case \ref{case:ftlr:move-cca-etc:splice-normal}:
+
+    We start by arguing the safety of $r_1$:
 

### 2018-06-28 fully-abstract-comp/ftlr-proof-cases.tex
Prove a result about xjmp.

--- 2018-06-27 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \item $\nonExec{\vsc_1'}$ and $\nonExec{\vsc_2'}$
-  \item $W_{R,1} \oplus W_{R,2}$ is defined
+  \item $W_{R,1} \oplus W_{R,2} \oplus W_M$ is defined
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
   \end{itemize}
@@
 \begin{proof}
-  TODO
+  From $\npair{(\sealed{\sigma,\vsc_1},\sealed{\sigma,\vsc_1'})} \in \lrvg{\trust}(W_{R,1})$, we know that
+  \begin{itemize}
+  \item $(\isLinear{\src{\vsc_1}} \text{ iff } \isLinear{\vsc_1'})$ 
+  \item $\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp$ 
+  \item $\pwheap[W_{R,1}](r) = (\pure,\_,H_\sigma)$ 
+  \item $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc$
+  \item $\npair[n']{\stpair[.]{\vsc_1}{\vsc_1'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,1})$ for all $n' < n$
+  \item If $(\isLinear{\src{\vsc_1}}$ then for all $W' \future W_{R,1}$, $W_o$, $n' < n$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+    \begin{equation*}
+      \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
+    \end{equation*}
+  \item If $(\nonLinear{\src{\vsc_1}}$ then for all $W' \future \purePart{W_{R,1}}$, $W_o$, $n' < n$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+    \begin{equation*}
+      \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
+    \end{equation*}
+  \end{itemize}
+
+  From $\npair{(\sealed{\sigma,\vsc_2},\sealed{\sigma,\vsc_2'})} \in \lrvg{\trust}(W_{R,2})$, we know that
+  \begin{itemize}
+  \item $(\isLinear{\src{\vsc_2}} \text{ iff } \isLinear{\vsc_2'})$ 
+  \item $\exists r' \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp$ 
+  \item $\pwheap[W_{R,2}](r') = (\pure,\_,H_\sigma)$ 
+  \item $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc$
+  \item $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$ for all $n' < n$
+  \end{itemize}
+
+  From $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$, we know that two different regions cannot have $H_\sigma$ defined for the same $\sigma$, so that $r = r'$.
+  Both when $\isLinear{\vsc_1}$ and when $\nonLinear{\vsc_1}$ (using Lemma~\ref{lem:world-fut-purePart}), we know that for all $W' \future W_{R,1}$, $W_o$, $n' < n$, $\vsc_2$, $\vsc_2'$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+  \begin{equation*}
+    \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
+  \end{equation*}
+  We can now instantiate this fact with $W' = W_{R,1}$, $W_o = W_{R,2}$, $n' = n-1$, and the fact that $\npair[n-1]{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$ (see above) to conclude
+  \begin{equation*}
+    \npair[n-1]{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W_{R,1}\oplus W_{R,2})\text{.}
+  \end{equation*}
 \end{proof}

### 2018-06-28 fully-abstract-comp/ftlr-proof-cases.tex
Oops, forgot a decrease in steps.
Prove FTLR xjmp case.

@@
   \begin{itemize}
-  \item $\npair{(\vsc_1,\vsc_2,\vsc_1',\vsc_2')} \in \lrexj(W_{R,1}\oplus W_{R,2})$
+  \item $\npair[n-1]{(\vsc_1,\vsc_2,\vsc_1',\vsc_2')} \in \lrexj(W_{R,1}\oplus W_{R,2})$
   \end{itemize}
@@
 
-  In the second case (xjmp), we have that $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trust}(W_{R,1})$, $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(W_{R,2})$ and $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(W_R')$ for some $W_{R,1}, W_{R,2}, W_R'$ with $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$.
-
-  Using the facts that $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$, and the above points, Lemma~\ref{lem:sealed-lrv-lrexj} tells us that $\npair{(c_1,c_2,c_1',c_2')} \in \lrexj(W_{R,1} \oplus W_{R,2})$.
-
-  TODO
+  In the second case (xjmp), we have that $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trust}(W_{R,1})$, $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(W_{R,2})$ and $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(\{r_1,r_2\})(W_R')$ for some $W_{R,1}, W_{R,2}, W_R'$ with $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$.
+
+  Using the facts that $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$, and the above points, Lemma~\ref{lem:sealed-lrv-lrexj} tells us that $\npair[n-1]{(c_1,c_2,c_1',c_2')} \in \lrexj(W_{R,1} \oplus W_{R,2})$.
+
+  By definition of $\lrexj$, it suffices to prove that
+  \begin{itemize}
+  \item
+    \begin{equation*}
+      \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
+      \lrr(\{\rdata\})(W_R')
+    \end{equation*}
+
+    It is easy to show that $\npair{\linCons{c_1},\linCons{c_1'}} \in \lrvg{\trust}(\purePart{W_{R,1}})$ and $\npair{\linCons{c_2},\linCons{c_2'}} \in \lrvg{\trust}(\purePart{W_{R,2}})$, by using Lemma~\ref{lem:non-linear-pure} in the non-linear case and the fact that $0$ is always related to itself by definition of $\lrvg{\trust}$ in the linear case.
+    We also have that $\purePart{W_R'} = \purePart{W_{R,1}} = \purePart{W_{R,2}}$ by Lemma~\ref{lem:purePart-oplus} and the fact that $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$, so it follows that $\npair{\linCons{c_1},\linCons{c_1'}} \in \lrvg{\trust}(\purePart{W_R'})$ and $\npair{\linCons{c_2},\linCons{c_2'}} \in \lrvg{\trust}(\purePart{W_R'})$.
+    From this and the fact that $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(\{r_1,r_2\})(W_R')$, it follows easily that
+    \begin{equation*}
+      \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
+      \lrr(\{\rdata\})(W_R')
+    \end{equation*}
+    as required.
+
+  \item $\memSat[n-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
+    Follows by Lemma~\ref{lem:downwards-closed} from the corresponding assumption of this lemma.
+
+  \item $W_M \oplus W_R' \oplus W_{R,1} \oplus W_{R,2}$ is defined.
+
+    This follows easily from the facts that $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$ (see above) and the assumption that $W_\pcreg \oplus W_R' \oplus W_M'$ is defined.
+  \end{itemize}
 \end{proof}

### 2018-06-28 fully-abstract-comp/ftlr-proof-cases.tex
Also forgot to use reasonability in the FTLR proof xjmp case, fixed now.

@@
   \]
+  If $n-1 = 0$, then this holds vacuously (by definition of $\lrol$ and $\lror$), so we assume that $n-1 > 0$.
 
@@
       \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
+      \lrrg{\trust}(\{\rdata\})(W_R')
+    \end{equation*}
+
+    From the fact that $\Phi_S$ is reasonable up to $n$ steps, tells us that $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n-1$ steps for all $r \neq \pcreg$.
+    Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} then tells us that
+    \begin{equation*}
+      \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
       \lrr(\{\rdata\})(W_R')
     \end{equation*}
-    as required.
+    using the fact that $n-1 > 0$ (by assumption above), Theorem~\ref{thm:ftlr} holds up to $n-1$ steps (by assumption).
 

### 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
Discharge two TODOs.
Proof of split lemma.

--- 2018-06-28 fully-abstract-comp/ftlr-proof-cases.tex
@@
   \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-  \item $b \le n \le e$
-  \item $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$
-  \item $c_2 = ((\perm,\lin),n,\eaddr,\aaddr)$
+  \item $b \le s < e$
+  \item $c_1 = ((\perm,\lin),\baddr,s,\aaddr)$
+  \item $c_2 = ((\perm,\lin),s+1,\eaddr,\aaddr)$
   \item $c_3 = \linCons{c}$
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W \oplus \purePart{W_1 \oplus W_2 \oplus W_3}}$
   \end{itemize}
@@
 \begin{proof}
-  TODO
-
-  From $\npair{(c,c)} \in \lrvg{\trust}(W)$, we get that either $\perm = \noperm$, in which case we can pick the ownership distribution to be the same as before, or $\npair{[\baddr,\eaddr]} \in \readCond{\linear,R_R(r_3)}$.
-
-    In the latter case, we get $S \subseteq \addressable{\linear,R_R(r_3)}$ and $R : S \fun \powerset{R_R(r_3)}$ such that $\forall r \in S \ldotp |R(r)| = 1$ and $\biguplus_{r\in S} R(r) \supseteq [\baddr,\eaddr]$.
-
-    This gives us a way to distribute the ownership, namely $r_3$ gives up ownership of regions $R^{-1}([\baddr,\eaddr])$, $r_1$ gets ownership of $R^{-1}([\baddr,n])$, and $r_2$ gets ownership over $R^{-1}([n+1,\eaddr])$.
-    The remainder of the ownership distribution remains unchanged.
-
-TODO TODO TODO     
-    Now we have to argue the following:
-    \begin{itemize}
-    \item if $\perm \in \readAllowed$, then
-      \begin{itemize}
-      \item $\npair{[\baddr,n]} \in \readCond{\linear,R_R'(r_1)}$
-      \item $\npair{[n+1,\eaddr]} \in \readCond{R_R'(r_2)}$
-      \end{itemize}
-
-    \item if $\perm \in \writeAllowed$, then
-      \begin{itemize}
-      \item $\npair{[\baddr,n]} \in \writeCond{\linear,R_R'(r_1)}$
-      \item $\npair{[n+1,\eaddr]} \in \writeCond{\linear,R_R'(r_2)}$
-      \end{itemize}
-    \item $\perm \neq \rx$
-
-      Assuming $\perm = \rx$, then it must have been the case that $\lin = \normal$ contradicting the assumption $\isLinear{(\Phi_S(r_3))}$.
-
-    \end{itemize}
-    or
-    \begin{itemize}
-    \item 
-    \end{itemize}
-    depending on whether the permission of the capability was $\perm = \noperm$
-    or not. In both cases, the result follows immediately by the assumptions.
-
-    If $\neg \isLinear{(\Phi_S(r_3))}$, then we pick $R_R' = R_R$
-
-    TODO
-
+
+  If $\lin = \normal$, then we pick $W_3 = W$ and $W_1 = W_2 = \purePart{W_3}$ which easily satisfies $W = W_1 \oplus W_2 \oplus W_3$.
+
+  Assuming $\gc = (\ta,\gc,\gsigrets,\gsigcloss)$, we know by assumption $\npair{(c,c)} \in \lrvg{\trust}(W)$ that either
+\begin{itemize}
+\item $\ta \subseteq [\baddr,\eaddr]$; or
+\item $\ta \# [\addr,\eaddr]$
+\end{itemize}
+In the former case, the result follows from Lemma~\ref{lem:non-linear-pure} and Lemma~\ref{lem:conds-shrinkable}.
+
+In the latter case, we know that $[\baddr,s],[s+1,\eaddr] \neq \emptyset$ because $b \le s < e$, so the result follows from Lemma~\ref{lem:non-linear-pure} and Lemma~\ref{lem:conds-shrinkable}.
+
+If $\lin = \linear$, then 
+we know from $\npair{(c,c)} \in \lrvg{\trust}(W)$ and w.l.o.g we can assume
+\begin{itemize}
+\item $\npair{[\baddr,\eaddr]} \in \readCond{\linear,W}$
+\item $\npair{[\baddr,\eaddr]} \in \writeCond{\linear,W}$
+\end{itemize}
+from this, we get $S_{\var{read}} \subseteq \addressable{\linear,\pwheap{W}}$ and
+$R_{\var{read}} : S_{\var{read}} \fun \Worlds$ such that
+\begin{itemize}
+\item $\biguplus_{r\in S_{\var{read}}} R_{\var{read}}(r) \supseteq A$
+\item $\forall r\ldotp |R_{\var{read}}(r)|  = 1$
+\item $\forall r \in S_{\var{read}} \ldotp \pwheap(r).H \nsubeq \stdreg{R_{\var{read}}(r),\gc}{\pur}.H$
+\end{itemize}
+and
+$S_{\var{write}} \subseteq \addressable{\linear,\pwheap{W}}$ and
+$R_{\var{write}} : S_{\var{write}} \fun \Worlds$ such that
+\begin{itemize}
+\item $\biguplus_{r\in S_{\var{write}}} R_{\var{write}}(r) \supseteq A$
+\item $\forall r\ldotp |R_{\var{write}}(r)|  = 1$
+\item $\forall r \in S_{\var{write}} \ldotp \pwheap(r).H \nsupeq
+  \stdreg{R_{\var{write}}(r),\gc}{\pur}.H \wedge \pwheap(r) \text{ is address-stratified}$
+\end{itemize}
+
+Now we would like to show $R_{\var{read}}^{-1}([\baddr,\eaddr]) = R_{\var{write}}^{-1}([\baddr,\eaddr])$.
+We know that the two sets are the same size as both $R$'s map to singleton sets.
+This means that there exists $r \neq r'$ s.t.\ $R_{\var{read}}(r) =
+R_{\var{write}}(r') = \{\aaddr'\}$ for $\aaddr' \in [\baddr,\eaddr]$.
+
+By definition of the standard region, we know that for any $\hat{W}$ and any $\ms$ and $\ms'$ where $\npair[n-1]{(\ms,\ms')} \in \pwheap(r)$ we have $\dom(\ms)= \dom(\ms') = \{a'\}$.
+
+By assumption $\pwheap(r)$ is address-stratified which means that for any $\hat{W}$ and any $\ms$ and $\ms'$ where $\npair[n-1]{(\ms,\ms')} \in \pwheap(r)$ we have $\dom(\ms)e= \dom(\ms') = \{a'\}$.
+
+By the memory satisfaction, the memory must be split into disjointed parts that
+each satisfy a region. With two regions that require a memory segment pair with
+the same domain, we cannot satisfy all the regions, so we must have
+
+Now pick $W_1$ as the world that owns $R_{\var{read}}^{-1}([\baddr,s])$, $W_2$
+the world that owns $R_{\var{read}}^{-1}([s+1,\eaddr])$, and $W_3$ as the world
+that owns the remaining regions of $W$.
+
+It is clearly the case that $W = W_1 \oplus W_2 \oplus W_3$.
+
+In this case, we need to show
+\[
+  \npair{(0,0)} \in \lrvg{\trust}(W_3)
+\]
+which is trivially the case.
+
+For the remaining, it suffices to show
+\begin{itemize}
+\item $\npair{[\baddr,s]} \in \readCond{\linear,W_1}$
+\item $\npair{[\baddr,s]} \in \writeCond{\linear,W_1}$
+\item $\npair{[s+1,\eaddr]} \in \readCond{\linear,W_2}$
+\item $\npair{[s+1,\eaddr]} \in \writeCond{\linear,W_2}$
+\end{itemize}
+which follows by assumption.
 \end{proof}
@@
   \item $\nonExec{\vsc_1'}$ and $\nonExec{\vsc_2'}$
-  \item $W_{R,1} \oplus W_{R,2} \oplus W_M$ is defined
-  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
+  \item $W_{R,1} \oplus W_{R,2}$ is defined
   \end{itemize}
@@
 \begin{proof}
-  From $\npair{(\sealed{\sigma,\vsc_1},\sealed{\sigma,\vsc_1'})} \in \lrvg{\trust}(W_{R,1})$, we know that
-  \begin{itemize}
-  \item $(\isLinear{\src{\vsc_1}} \text{ iff } \isLinear{\vsc_1'})$ 
-  \item $\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp$ 
-  \item $\pwheap[W_{R,1}](r) = (\pure,\_,H_\sigma)$ 
-  \item $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc$
-  \item $\npair[n']{\stpair[.]{\vsc_1}{\vsc_1'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,1})$ for all $n' < n$
-  \item If $(\isLinear{\src{\vsc_1}}$ then for all $W' \future W_{R,1}$, $W_o$, $n' < n$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
-    \begin{equation*}
-      \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
-    \end{equation*}
-  \item If $(\nonLinear{\src{\vsc_1}}$ then for all $W' \future \purePart{W_{R,1}}$, $W_o$, $n' < n$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
-    \begin{equation*}
-      \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
-    \end{equation*}
-  \end{itemize}
-
-  From $\npair{(\sealed{\sigma,\vsc_2},\sealed{\sigma,\vsc_2'})} \in \lrvg{\trust}(W_{R,2})$, we know that
-  \begin{itemize}
-  \item $(\isLinear{\src{\vsc_2}} \text{ iff } \isLinear{\vsc_2'})$ 
-  \item $\exists r' \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp$ 
-  \item $\pwheap[W_{R,2}](r') = (\pure,\_,H_\sigma)$ 
-  \item $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc$
-  \item $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$ for all $n' < n$
-  \end{itemize}
-
-  From $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$, we know that two different regions cannot have $H_\sigma$ defined for the same $\sigma$, so that $r = r'$.
-  Both when $\isLinear{\vsc_1}$ and when $\nonLinear{\vsc_1}$ (using Lemma~\ref{lem:world-fut-purePart}), we know that for all $W' \future W_{R,1}$, $W_o$, $n' < n$, $\vsc_2$, $\vsc_2'$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
-  \begin{equation*}
-    \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
-  \end{equation*}
-  We can now instantiate this fact with $W' = W_{R,1}$, $W_o = W_{R,2}$, $n' = n-1$, and the fact that $\npair[n-1]{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$ (see above) to conclude
-  \begin{equation*}
-    \npair[n-1]{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W_{R,1}\oplus W_{R,2})\text{.}
-  \end{equation*}
+  TODO
 \end{proof}

### 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex

@@
   \begin{itemize}
-  \item $\npair{(\vsc_1,\vsc_2,\vsc_1',\vsc_2')} \in \lrexj(W_{R,1}\oplus W_{R,2})$
-  \end{itemize}
-\end{lemma}
-\begin{proof}
-  TODO
+  \item $\npair[n-1]{(\vsc_1,\vsc_2,\vsc_1',\vsc_2')} \in \lrexj(W_{R,1}\oplus W_{R,2})$
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  From $\npair{(\sealed{\sigma,\vsc_1},\sealed{\sigma,\vsc_1'})} \in \lrvg{\trust}(W_{R,1})$, we know that
+  \begin{itemize}
+  \item $(\isLinear{\src{\vsc_1}} \text{ iff } \isLinear{\vsc_1'})$ 
+  \item $\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp$ 
+  \item $\pwheap[W_{R,1}](r) = (\pure,\_,H_\sigma)$ 
+  \item $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc$
+  \item $\npair[n']{\stpair[.]{\vsc_1}{\vsc_1'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,1})$ for all $n' < n$
+  \item If $(\isLinear{\src{\vsc_1}}$ then for all $W' \future W_{R,1}$, $W_o$, $n' < n$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+    \begin{equation*}
+      \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
+    \end{equation*}
+  \item If $(\nonLinear{\src{\vsc_1}}$ then for all $W' \future \purePart{W_{R,1}}$, $W_o$, $n' < n$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+    \begin{equation*}
+      \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
+    \end{equation*}
+  \end{itemize}
+
+  From $\npair{(\sealed{\sigma,\vsc_2},\sealed{\sigma,\vsc_2'})} \in \lrvg{\trust}(W_{R,2})$, we know that
+  \begin{itemize}
+  \item $(\isLinear{\src{\vsc_2}} \text{ iff } \isLinear{\vsc_2'})$ 
+  \item $\exists r' \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp$ 
+  \item $\pwheap[W_{R,2}](r') = (\pure,\_,H_\sigma)$ 
+  \item $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc$
+  \item $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$ for all $n' < n$
+  \end{itemize}
+
+  From $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$, we know that two different regions cannot have $H_\sigma$ defined for the same $\sigma$, so that $r = r'$.
+  Both when $\isLinear{\vsc_1}$ and when $\nonLinear{\vsc_1}$ (using Lemma~\ref{lem:world-fut-purePart}), we know that for all $W' \future W_{R,1}$, $W_o$, $n' < n$, $\vsc_2$, $\vsc_2'$, $\npair[n']{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+  \begin{equation*}
+    \npair[n']{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W'\oplus W_o))
+  \end{equation*}
+  We can now instantiate this fact with $W' = W_{R,1}$, $W_o = W_{R,2}$, $n' = n-1$, and the fact that $\npair[n-1]{\stpair[.]{\vsc_2}{\vsc_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$ (see above) to conclude
+  \begin{equation*}
+    \npair[n-1]{\src{\vsc_1},\src{\vsc_2},\vsc_1',\vsc_2'} \in \lrexj(W_{R,1}\oplus W_{R,2})\text{.}
+  \end{equation*}
 \end{proof}
@@
 
-  In the second case (xjmp), we have that $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trust}(W_{R,1})$, $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(W_{R,2})$ and $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(W_R')$ for some $W_{R,1}, W_{R,2}, W_R'$ with $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$.
-
-  Using the facts that $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$, and the above points, Lemma~\ref{lem:sealed-lrv-lrexj} tells us that $\npair{(c_1,c_2,c_1',c_2')} \in \lrexj(W_{R,1} \oplus W_{R,2})$.
-
-  TODO
+  In the second case (xjmp), we have that $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trust}(W_{R,1})$, $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trust}(W_{R,2})$ and $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(\{r_1,r_2\})(W_R')$ for some $W_{R,1}, W_{R,2}, W_R'$ with $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$.
+
+  Using the facts that $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$, and the above points, Lemma~\ref{lem:sealed-lrv-lrexj} tells us that $\npair[n-1]{(c_1,c_2,c_1',c_2')} \in \lrexj(W_{R,1} \oplus W_{R,2})$.
+
+  By definition of $\lrexj$, it suffices to prove that
+  \begin{itemize}
+  \item
+    \begin{equation*}
+      \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
+      \lrr(\{\rdata\})(W_R')
+    \end{equation*}
+
+    It is easy to show that $\npair{\linCons{c_1},\linCons{c_1'}} \in \lrvg{\trust}(\purePart{W_{R,1}})$ and $\npair{\linCons{c_2},\linCons{c_2'}} \in \lrvg{\trust}(\purePart{W_{R,2}})$, by using Lemma~\ref{lem:non-linear-pure} in the non-linear case and the fact that $0$ is always related to itself by definition of $\lrvg{\trust}$ in the linear case.
+    We also have that $\purePart{W_R'} = \purePart{W_{R,1}} = \purePart{W_{R,2}}$ by Lemma~\ref{lem:purePart-oplus} and the fact that $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$, so it follows that $\npair{\linCons{c_1},\linCons{c_1'}} \in \lrvg{\trust}(\purePart{W_R'})$ and $\npair{\linCons{c_2},\linCons{c_2'}} \in \lrvg{\trust}(\purePart{W_R'})$.
+    From this and the fact that $\npair{(\Phi_S'',\Phi_T'')} \in \lrrg{\trust}(\{r_1,r_2\})(W_R')$, it follows easily that
+    \begin{equation*}
+      \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
+      \lrrg{\trust}(\{\rdata\})(W_R')
+    \end{equation*}
+
+    From the fact that $\Phi_S$ is reasonable up to $n$ steps, tells us that $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n-1$ steps for all $r \neq \pcreg$.
+    Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} then tells us that
+    \begin{equation*}
+      \npair[n-1]{\left(\array{l}\Phi_S.\reg\update{r_1,r_2}{\linCons{c_1},\linCons{c_2}},\\ \Phi_T.\reg\update{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}\endarray\right)} \in
+      \lrr(\{\rdata\})(W_R')
+    \end{equation*}
+    using the fact that $n-1 > 0$ (by assumption above), Theorem~\ref{thm:ftlr} holds up to $n-1$ steps (by assumption).
+
+  \item $\memSat[n-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
+    Follows by Lemma~\ref{lem:downwards-closed} from the corresponding assumption of this lemma.
+
+  \item $W_M \oplus W_R' \oplus W_{R,1} \oplus W_{R,2}$ is defined.
+
+    This follows easily from the facts that $W_R = W_{R,1} \oplus W_{R,2} \oplus W_R'$ (see above) and the assumption that $W_\pcreg \oplus W_R' \oplus W_M'$ is defined.
+  \end{itemize}
 \end{proof}
@@
 \begin{proof}
-  TODO
+  Follows by inspecting the definition of $\memSatFStack{\ms_\stk,\ms_T}{W}$ and by the monotonicity of memory relations in the world.
 \end{proof}
@@
 \begin{proof}
-  TODO
+  Follows by inspecting the definition of $\lrheap$ and $\purePart{}$ and using the monotonicity of memory relations in the world.
 \end{proof}

### 2018-06-29 fully-abstract-comp/technical-report.tex
Prove the jmp/jnz case of the FTLR.

@@
       \begin{array}{r l l}
+        \perm \not\in \{\rx,\rwx\} \wedge\\
         % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
@@
     \item $r_1 \neq \rtmp{1}$
-    \item $\Phi(\pcreg) + \calllen$ behaves reasonably up to $n'-1$ steps
+    \item $n' = 0$ or $\Phi(\pcreg) + \calllen$ behaves reasonably up to $n'-1$ steps
     \end{itemize}
@@
       \begin{itemize}
-      \item $\exec{\Phi(r_2)}$ and $\Phi(r_2)$ behaves reasonably up to $n' - 1$ steps.
-      \item $\nonExec{\Phi(r_2)}$ and $\Phi(r_2)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$.
+      \item $\exec{\Phi(r_2)}$ and $n' = 0$ or $\Phi(r_2)$ behaves reasonably up to $n' - 1$ steps.
+      \item $\nonExec{\Phi(r_2)}$ and $n' = 0$ or $\Phi(r_2)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$.
       \end{itemize}
@@
     \end{itemize}
-    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ step
+    Then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ step
   \item \emph{Don't leak private stuff...} If
@@
         ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-      \item $\Phi'$ is reasonable up to $n'-1$ steps
+      \item $\Phi$ does not point to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+      \item $\Phi$ does not point to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$, $\offpc$, $\offsigma$
+      \item $n' = 0$ or $\Phi'$ is reasonable up to $n'-1$ steps
       \end{itemize}
@@
       \item $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
-      \item $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
+      \item $n' = 0$ or $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
       \end{itemize}
@@
       \begin{itemize}
-      \item $\Phi$ points to $\src{\txjmp{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
-      \item $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
+      \item $\Phi$ points to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+      \item $n' = 0$ or $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
       \end{itemize}

### 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
Reply to Dominique comment.

@@
   The property for $\lrv$ follows from the other properties, by definition, from Lemma~\ref{lem:n-equality-props}, non-expansiveness of regions \dominique{Is $\xi$ non-expansive/contractive/?}
+  \lau{I believe it is at-least non-expansive and monotone as it is a morphism in the category of preordered c.o.f.e.'s.
+    Do you also ``need'' it to be contractive?}
   TODO

### 2018-06-29 fully-abstract-comp/technical-report.tex
Remove old definition of reasonability.

@@
 
-\subsection{Alternative formulation of reasonability?}
+\subsection{Reasonable Components}
 \label{sec:reasonability}
@@
 
-\subsection{Reasonable components}
-\label{sec:reas-comp}
-\begin{definition}
-  \label{def:points-to-instr}
-  A configuration $(\ms,\reg)$ {\sourcecolor or $\Phi = (\ms,\reg,\_,\_)$} is said to point to non-$\scall{}{}{}$ instruction $i$ in $A$ iff
-  \[
-    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand \aaddr \in A
-  \]
-and
-\[
-  \decInstr{\ms(\aaddr)} = i
-\]
-\end{definition}
-
-\begin{definition}
-  \label{def:points-to-call}
-  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{\offpc,\offsigma}{r_1}{r_2}$ in $A$ iff
-  \[
-    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand [\aaddr,\aaddr+\calllen-1] \subseteq A
-  \]
-and
-\[
-  \aaddr+\calllen - 1 \leq \eaddr
-\]
-and
-\[
-  \callCond{\Phi,r_1,r_2,\offpc,\offsigma,\aaddr}
-\]
-\end{definition}
-We want to make a conditional full-abstraction statement.
-The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do.
-(2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation.
-\begin{definition}
-  \label{def:check-stack-addr-before-call}
-  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ iff\\
-  for all $i$ if
-  \[
-    \src{\Phi} \nstep[i]{(\ta,\stkb)} \src{\Phi'}
-  \]
-  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
-  % and there exists no $j<i$ s.t.
-  % \[
-  %   \src{\Phi} \nstep[j]{} \src{\Phi''}
-  % \]
-  % where $\src{\Phi''}$ points to $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
-  then
-  \[
-    \src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
-  \]
-  and
-  \[
-    r_1 \neq \rtmp{1}
-  \]
-  \lau{maybe the last bit should have its own definition.}
-  % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
-\end{definition}
-
-\begin{definition}[Use return seals only for call]
-  \label{def:use-return-seals-call}
-  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
-  for all $i$, if
-  \[
-    \src{\Phi} \nstep[i]{(\ta,\_)} \src{\Phi'}
-  \]
-  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either ($\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$) or $\sigma \in \sigcloss$.
-\end{definition}
-
-\begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
-  \label{def:handle-trusted-seals-properly}
-  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
-  for all $i$, if
-  \[
-    \src{\Phi} \nstep[i]{(\ta,\stkb)} \src{\Phi'} \step[(\ta,\stkb)] \src{\Phi''}
-  \]
-  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
-  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
-  \[
-    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow (\sigrets \cup \sigcloss) \mathrel{\#} [\sigma_b,\sigma_e]
-  \]
-and
-%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
-  \[
-    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
-  \]
-  and
-    \begin{multline*}
-      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in (\sigrets\cup\sigcloss) \Rightarrow 
-      \\  \left(  
-      \begin{array}{r l}
-        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow (\sigrets\cup\sigcloss) \# [\sigma_b,\sigma_e]) \wedge\\
-        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
-      \end{array}
-      \right)
-  \end{multline*}
-\end{definition}
-
-%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}
-\begin{definition}[Never store link seals and code capability]
-  \label{def:never-store-seal-code-cap}
-  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ iff\\
-  for all $i$, if
-  \[
-    \src{\Phi} \nstep[i]{(\ta,\_)} \src{\Phi'}
-  \]
-  and $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
-\[
-  \Phi'(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
-\]
-and
-\[
-  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
-\]
-\end{definition}
-
-%   \dominique{TODO: Also not allowed: sealing such a value with a seal that you do not own.}
-%   \lau{is this covered by definition 14 (Use return seals only for call)? }
-% \begin{definition}[Never seal link seals and code capabilities with unknown seal]  
-% \end{definition}
-
-% \dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures).
-That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.
-Update: this is resolved}
-\begin{definition}
-  A component $\var{comp}$ with $\dom(\var{comp}.\mscode) = A$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\var{comp}}$ satisfies:
-  \begin{itemize}
-  \item $\plug{\context}{\var{comp}}$ checks $\stkb$ before calls in $A$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
-  \item $\plug{\context}{\var{comp}}$ uses link seals properly in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
-  \item $\plug{\context}{\var{comp}}$ handles trusted seals properly in $A$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
-  \item $\plug{\context}{\var{comp}}$ does not store link seals and code capabilities in $A$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
-  \end{itemize}
-\end{definition}
-
-\subsection{Full abstraction: precise formulation}
-\label{sec:full-abstr-prec}
-
 \begin{definition}[Source language contextual equivalence]

### 2018-06-29 fully-abstract-comp/popl2019/paper.tex
First sketch of POPL submission.

--- 2017-12-07 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex
@@
-\documentclass[sigplan,screen]{acmart}
+%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For double-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
+%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For single-blind review submission, w/ CCS and ACM Reference
+%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%% For final camera-ready submission, w/ required CCS and ACM Reference
+%\documentclass[acmsmall]{acmart}\settopmatter{}
 
-\usepackage{booktabs} % For formal tables
-\usepackage{cleveref}
-\usepackage{hyperref}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
+%% Journal information
+%% Supplied to authors by publisher for camera-ready submission;
+%% use defaults for review submission.
+% \acmJournal{PACMPL}
+% \acmVolume{1}
+% \acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
+% \acmArticle{1}
+% \acmYear{2018}
+% \acmMonth{1}
+% \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+% \startPage{1}
 
-% Copyright
-\setcopyright{rightsretained}
-\acmDOI{}
-\acmISBN{}
-\acmConference[PriSC]{Workshop on Principles of Secure Compilation}{2018}{Los Angeles} 
-\acmYear{2018}
-\copyrightyear{2018}
-\settopmatter{printacmref=false}
+%% Copyright information
+%% Supplied to authors (based on authors' rights management selection;
+%% see authors.acm.org) by publisher for camera-ready submission;
+%% use 'none' for review submission.
+\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\copyrightyear{2018}           %% If different from \acmYear
 
-% \acmPrice{15.00}
+%% Bibliography style
+\bibliographystyle{ACM-Reference-Format}
+%% Citation style
+%% Note: author/year citations are required for papers published as an
+%% issue of PACMPL.
+\citestyle{acmauthoryear}   %% For author/year citations
 
-%\acmBadgeL[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
-%\acmBadgeR[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
+% %% Some recommended packages.
+% \usepackage{booktabs}   %% For formal tables:
+%                         %% http://ctan.org/pkg/booktabs
+% \usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+%                         %% http://ctan.org/pkg/subcaption
 
 \begin{document}
+
+%% Title information
 \title{Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
-% \titlenote{Produces the permission block, and
-%   copyright information}
-\subtitle{Extended Abstract}
-% \subtitlenote{The full version of the author's guide is available as
-%   \texttt{acDat zou in dmart.pdf} document}
+% \titlenote{with title note}
+% \subtitle{Fully abstract overlay semantics}
+% \subtitlenote{with subtitle note}
+
+%% Author information
+%% Contents and number of authors suppressed with 'anonymous'.
+%% Each author should be introduced by \author, followed by
+%% \authornote (optional), \orcid (optional), \affiliation, and
+%% \email.
+%% An author may have multiple affiliations and/or emails; repeat the
+%% appropriate command.
+%% Many elements are not rendered, but should be provided for metadata
+%% extraction tools.
 
 \author{Lau Skorstengaard}
-% \orcid{}
-\affiliation{%
+% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
   \institution{Aarhus University}
+  \country{Denmark}                    %% \country is recommended
 }
-\email{lau@cs.au.dk}
+\email{lask@cs.au.dk}
 
+%% Author with two affiliations and emails.
 \author{Dominique Devriese}
-\orcid{0000-0002-3862-6856}
-\affiliation{%
-  \institution{imec-DistriNet, KU Leuven}
+\orcid{0000-0002-3862-6856}             %% \orcid is optional
+\affiliation{
+  \institution{KU~Leuven}           %% \institution is required
+  \country{Belgium}                   %% \country is recommended
 }
-\email{dominique.devriese@cs.kuleuven.be}
+\email{dominique.devriese@cs.kuleuven.be}         %% \email is recommended
 
 \author{Lars Birkedal}
-\affiliation{%
+% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
   \institution{Aarhus University}
+  \country{Denmark}                    %% \country is recommended
 }
 \email{birkedal@cs.au.dk}
-
-% The default list of authors is too long for headers}
-% \renewcommand{\shortauthors}{B. Trovato et al.}
 
@@
 The calling convention is based on linear capabilities, a type of capabilities that are prevented by the hardware from being duplicated.
-In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, building on the concept of fully abstract compilation.
+In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, using what we call a fully abstract overlay semantics.
 \end{abstract}
 
-%
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
-%
+% %% 2012 ACM Computing Classification System (CSS) concepts
+% %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
 % \begin{CCSXML}
@@
 % <concept>
-%   <concept_id>10010520.10010553.10010562</concept_id>
-%   <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+% <concept_id>10011007.10011006.10011008</concept_id>
+% <concept_desc>Software and its engineering~General programming languages</concept_desc>
 % <concept_significance>500</concept_significance>
@@
 % <concept>
-%   <concept_id>10010520.10010575.10010755</concept_id>
-%   <concept_desc>Computer systems organization~Redundancy</concept_desc>
+% <concept_id>10003456.10003457.10003521.10003525</concept_id>
+% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
 % <concept_significance>300</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10010520.10010553.10010554</concept_id>
-%   <concept_desc>Computer systems organization~Robotics</concept_desc>
-%   <concept_significance>100</concept_significance>
-%  </concept>
-%  <concept>
-%   <concept_id>10003033.10003083.10003095</concept_id>
-%   <concept_desc>Networks~Network reliability</concept_desc>
-%   <concept_significance>100</concept_significance>
 % </concept>
@@
 
-% \ccsdesc[500]{Computer systems organization~Embedded systems}
-% \ccsdesc[300]{Computer systems organization~Redundancy}
-% \ccsdesc{Computer systems organization~Robotics}
-% \ccsdesc[100]{Networks~Network reliability}
+% \ccsdesc[500]{Software and its engineering~General programming languages}
+% \ccsdesc[300]{Social and professional topics~History of programming languages}
+%% End of generated code
 
-% \keywords{ACM proceedings, \LaTeX, text tagging}
-
-% \begin{teaserfigure}
-%   \includegraphics[width=\textwidth]{sampleteaser}
-%   \caption{This is a teaser}
-%   \label{fig:teaser}
-% \end{teaserfigure}
+\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, overlay semantics}
 
@@
 
-\section{Linear Capabilities, not Local}
-What we essentially want to achieve in a secure calling convention is to give an invoked function access to a stack pointer and a return pointer, but only temporarily: during the execution of the invoked function.
-In our original approach, we make both pointers local, so that the CHERI processor guarantees, essentially, that they can only be stored in registers or on the stack.
-To prevent an invoked function from transferring the pointers across outside the lifetime of its stack frame, it then suffices to block only these two communication channels, rather than the entire memory.
-In other words, we need to clear registers and the stack before passing control to untrusted code, to prevent it from obtaining access to stack and return pointers for other stack frames than its own.
-The requirement for clearing registers is unproblematic, but clearing the entire stack upon every boundary crossing is a non-trivial requirement.
-We speculate that it can be made efficient with custom hardware extensions\footnote{Specifically, we imagine a kind of pre-L0 processor cache that does not retain values but only remembers that a certain range has been zeroed out.}, but this is not obvious.
+% %% Acknowledgments
+% \begin{acks}                            %% acks environment is optional
+%                                         %% contents suppressed with 'anonymous'
+%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+%   %% acknowledge financial support and will be used by metadata
+%   %% extraction tools.
+%   This material is based upon work supported by the
+%   \grantsponsor{GS100000001}{National Science
+%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+%   conclusions or recommendations expressed in this material are those
+%   of the author and do not necessarily reflect the views of the
+%   National Science Foundation.
+% \end{acks}
 
-An alternative that does not have this requirement, relies on \emph{linear capabilities}, rather than local ones.
-Although details in the literature are scarce, linear capabilities have already been implemented in the SAFE machine~\citep{amorim_verified_2014}\footnote{Private communication with Catalin Hrit\c{c}u.} and considered for implementation in CHERI\footnote{Private communication with Robert Watson.}.
-The idea is that all capabilities are marked as either regular or linear.
-Linear ones are subject to special hardware-enforced restrictions, that prevent them from being duplicated.
-For example, when linear capabilities are moved to a new location in registers or memory, they are erased from their old location.
-A few special instructions are needed to make this workable, e.g.\ a split instruction that splits a memory capability for a region of memory into disjoint capabilities for two parts of the region and a splice instruction that does the reverse.
-This linearity is a powerful tool: when passing a linear capability to untrusted code and receiving the same capability back, we are sure that they cannot have stored a copy of it.
-Similarly, when we receive, from untrusted code, a linear capability that addresses memory that we know is only ever addressed by linear capabilities, then we are sure that the untrusted code no longer has access that memory.
-
-As such, linear capabilities can be used as an alternative to local capabilities to enforce well-bracketed control flow.
-We cannot provide full details here for space reasons, but the idea is essentially the following.
-By passing stack and return pointers as linear capabilities, and requiring that invoked functions hand in their stack pointer upon return, untrusted code is prevented from holding on to stack and return pointers past their intended lifetime, essentially because they can only ever have one copy and they have to hand it in when this lifetime ends.
-
-\section{Formalising well-bracketed control flow}
-The second improvement we are making is related to the properties of well-bracketed control flow and local state encapsulation.
-To establish these properties, our previous work followed previous work \citep{dreyer_impact_2010} in providing a sound reasoning technique and using it to prove soundness of challenging example programs that rely on the properties.
-In the current work, we are investigating an alternative way to formalise the two properties, based on the notion of fully abstract compilation \citep{abadi_protection_1999}.
-
-We intend to start by formalising the assembly language for a capability machine with linear capabilities (the target language).
-Then, we will introduce the source language: a variant of this target language that features a primitive stack, built into the operational semantics, as well as additional instructions like call.
-While the target language is a standard assembly language without even the notion of a stack, the source language does have a stack, and control flow well-bracketedness and local state encapsulation follow more directly from the operational semantics.
-We will formalise our calling convention in the definition of a compiler from the source to the target language.
-Additionally, we conjecture that this compiler can be proved fully abstract, thus formally expressing and establishing security of the calling convention.
-
-Interestingly, the proof of full abstraction for this compiler is expected to use a very simple back-translation, that simply maps every target language instruction to the corresponding instruction in the source.
-This back-translation works because we construct our source language in such a way that it has direct counterparts to every target language value.
-Specifically, the source language contains a representation of stack and return pointers as unspecified abstract values whose behavior corresponds to their target counterparts, but in a way that guarantees the intended properties of well-bracketed control flow and local state encapsulation.
-
-\bibliographystyle{ACM-Reference-Format}
+\section{Linear Stack and Return Capabilities}
+\begin{itemize}
+\item informally explain how an adversary may try to abuse stack and return pointers
+\item informally explain how we prevent this using linear capabilities
+\item use the tikz pictures from the PriSC presentation to explain all of this
+\end{itemize}
+\section{A Capability Machine with Sealing and Linear Capabilities}
+\begin{itemize}
+\item present our target language and its operational semantics (excerpts)
+\item mention roughly what components look like
+\end{itemize}
+\section{Formulating Security with a Fully Abstract Overlay Semantics}
+\begin{itemize}
+\item present our source language, its operational semantics (excerpts)
+\item mention our assumption of reasonability
+\item present the full abstraction theorem.
+\item sketch high-level structure of the proof
+\end{itemize}
+\section{Discussion}
+\begin{itemize}
+\item mention that tail calls are supported through xcall
+\item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
+\end{itemize}
+\section{Related Work}
+\begin{itemize}
+\item CFI and friends
+\item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
+\item undefined behavior and full abstraction \citep{juglaret_beyond_2016}
+\end{itemize}
 \bibliography{references}
 
+%% Appendix
+% \appendix
+% \section{Appendix}
+
+% Text of appendix \ldots
+
 \end{document}

### 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
Fix a label.
Checked and removed TODO

@@
   \item Case \ref{case:ftlr:move-cca-etc:move}:
-    TODO: check if this case considers the trusted case.
-
     Pick the ownership distribution based on the linearity of $w_2$: If $\isLinear{w_2}$, then pick

### 2018-06-29 fully-abstract-comp/popl2019/paper.tex
Some more explanation in the intro about our contributions.

@@
 The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
-More suitable security primitives are offered by a type of processors known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
+More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
 These processors use tagged memory to enforce a strict separation between integers and \emph{capabilities}: pointers that carry authority.
 Capabilities come in different flavours.
-Memory capabilities are used to read from and write to memory.
-Additionally, the CHERI capability machine offers sealed code-data pairs of capabilities that represent an encapsulated closure: a piece of code coupled with private state that it gains access to upon invocation.\footnote{Other capability machines offer other mechanisms that can be used to the same effect.}
+Memory capabilities allow reading from and writing to a block of memory.
+Additionally, capability machines offer some form of \emph{object capabilities} that represent low-level encapsulated closures, i.e.
+a piece of code coupled with private state that it gains access to upon invocation.
+The concrete mechanics of object capabilities varies between different capability machines.
+On a recent capability machine called CHERI, for example, they take the form of pairs of capabilities sealed with a common seal that represent the code and data parts of the closure.
+The pair is opaque, but is transparently unsealed by the hardware upon invocation~\citep{watson_capability_2015}.
 
-CheriBSD, an operating system built on CHERI, uses per-component separate stacks and a central, trusted stack manager component to enforce well-bracketed control flow and private state encapsulation \citep{watson_cheri:_2015}.
-To prevent the stack pointer (or derivatives) from being passed to other components, CheriBSD uses local capabilities: a type of capabilities that can be kept in registers but not stored in memory (except through store-local capabilities like the stack pointer itself).
-The literature does not contain many details about the mechanism, let alone a formal analysis, but additionally, the mechanism does not seem able to support higher-order interfaces (e.g.
-C function pointers) in a scalable fashion.
+To enforce well-bracketed control flow and local state encapsulation on a capability machine, there are essentially two approaches.
+A first approach uses separate stacks for distrusting components, and a central, trusted stack manager component that mediates cross-component invocations.
+This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
+First, it scales poorly to large amounts of distrusting components, because of the need to reserve separate stack space for all components.
+Also, in the presence of higher-order values (e.g.
+function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to, in order to provide it the right stack pointer upon invocation and it is not clear how this can be done efficiently in the presence of large amounts of components.
+Finally, this approach does not allow passing stack references between components.
 
-In recent work, we have been looking at an alternative approach \citep{skorstengaard_reasoning_2017}.
-We also use CHERI's local capabilities, but combine them with a more traditional, single, shared stack that scales better in the presence of higher-order interfaces.
-By passing stack and return pointers as local capabilities, we effectively prevent callees from storing them on the heap.
-Using a step-indexed Kripke logical relation, we are able to prove correctness of programs that rely on well-bracketed control flow in a complex way.
+A more scalable approach retains a single stack that is shared between components.
+Enforce well-bracketed control flow and stack frame encapsulation in this approach, requires a way to temporarily provide stack and return capabilities to an untrusted component and revoke them after it returns.
+While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
+For example, CHERI offers a form of \emph{local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
+In previous work, we have demonstrated that by making the stack and return pointer local, and a number of security checks and measures, well-bracketedness and local stack frame encapsulation can be guaranteed~\citep{skorstengaard_reasoning_2017}.
+However, a problem with this approach is that revoking the local stack and return capabilities on every boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
 
-While this approach works quite well, this paper investigates another alternative approach.
-We aim to make two improvements, discussed in the next sections: (1) remove the need to clear the stack on boundary crossings and (2) explore a new formal statement of well-bracketed control flow and local state encapsulation.
+In this work, we propose and study another way to enforce well-bracketed control flow and local state encapsulation with a single shared stack.
+Instead of CHERI's local capabilities, we build on \emph{linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
+These capabilities are enforced by the hardware to be non-duplicable.
+We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
+The non-duplicability of linear capabilities, together with some security checks, then allows us to guarantee well-bracketed control flow and local stack frame encapsulation.
 
-% %% Acknowledgments
+A second contribution of this work is the way in which we formulate these two properties, using a technique we call \emph{fully abstract overlay semantics}.
+Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
+Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
+To accomodate this, we define a second operational semantics for our target language, which has a native well-bracketed stack and primitive ways to do calls and returns.
+This well-behaved semantics natively guarantees well-bracketed control flow and local stack frame encapsulation for components using our calling convention.
+As such, these components can be sure that they will only ever interact with well-behaved other components that respect our desirable properties.
+To express security of our calling convention, we then show that considering the same components in the original semantics, does not give adversaries additional ways to interact with them. 
+More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.
+that components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
+
+This approach expresses what it means to enforce our desirable properties, in a general way and makes it clear that we can support a very general class of programs.
+Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
+The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
+This independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
+
 % \begin{acks}                            %% acks environment is optional

### 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
Referencing issue
Updated seal instruction to actually check against seal range.
Do not forget to use reasonability in the store case...

--- 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
@@
           \sigma_{\eaddr,3}, \sigma}$
-      \item \label{case:ftlr:jnz-z} (jnz zero case) $k = 0$
+      \item \label{case:ftlr:move-cca-etc:jnz-z} (jnz zero case) $k = 0$
       \end{enumproof}
@@
 
-    Follows from Lemma~\ref{lem:store-reg-works}, redistributing ownership in the obvious way.
+    From $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we know that $\npair{(w,w')} \in \lrvg{\trust}(W_{R,2})$, $\npair{(w_2,w_2')} \in \lrvg{\trust}(\purePart{W_{R,2}]})$ and $\npair{(\Phi_S.\reg\update{r_2}{w_2},\Phi_T.\reg\update{r_2}{w_2'})} \in \lrrg{\trust}(W_R')$ with $W_R =W_R' \oplus W_{R,2}$ (using Lemma~\ref{lem:purePart-oplus} and~\ref{lem:purePart-duplicable}).
+
+    From reasonability of $\Phi_S$, we know that $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n-1$ steps.
+    Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} then tells us that $\npair{(w,w')} \in \lrv(W_{R,2})$.
+
+    The result then follows from Lemma~\ref{lem:store-reg-works}, redistributing ownership in the obvious way.
 

### 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
Fix use of reasonability in the call case.

@@
 
-  \lau{TODO: look through this case and find the place where the ''no seal cross boundary thing'' should have been used.
-    (I have done this, and it does not seem necessary (?))}
   In case \ref{case:ftlr:call}, we may assume $r_1 \neq r_2$\footnote{If the register contains a data capability, then the execution fails in the step after the jump.
@@
 
-  Case \ref{case:ftlr:scall:reg}: We need to split the ownership of $W_R'$.
-  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we get a way to split the ownership of $W_R$.
+  Case \ref{case:ftlr:scall:reg}:
+
+  First, from reasonability of $\Phi_S$, we get that $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n-1$ steps for all $r \neq \pcreg$.
+  Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} then tells us that
+  $\npair[n-1]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrr(W_R)$.
+
+  We then need to split the ownership of $W_R'$.
+  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
   We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership (i.e.\ $\purePart{W_{R,1}'}$).
@@
 
-  we split the world in the same way for the registers that remain unchanged, the
-  result follows from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity}.
-  \dominique{No: we need to use Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} together with the reasonability of registers at a boundary crossing (if there is one).}
-
-  This leaves the following cases:
+  we split the world in the same way for the registers that remain unchanged, and we get from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity} that
+  $\npair[n-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrrg{\trust}(\{\rdata,\rstk,\rtmp{1},r_1,r_2,\rretd,\rretc\})(W_R'')$ for the appropriate $W_R''$.
+
+  To obtain $\npair[n-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$, it remains to prove the following cases:
   \begin{description}

### 2018-06-30 fully-abstract-comp/ftlr-proof-cases.tex
Discharge a minor TODO.

@@
   \]
-  which gives us $R_R : (\RegName \setminus \{\pcreg\}) \fun World$ such that $W_R = \bigoplus_{r\in \RegName \setminus \{\pcreg\}} R_R(r)$ and for all $r$ in $\RegName \setminus \{\pcreg\}$ we have $\npair[n]{(\Phi_S.\reg(r),\Phi_T.\reg(r))} \in \lrrg{\trust}(R_R(r))$.
+  which gives us $R_R : (\RegName \setminus \{\pcreg\}) \fun \World$ such that $W_R = \bigoplus_{r\in \RegName \setminus \{\pcreg\}} R_R(r)$ and for all $r$ in $\RegName \setminus \{\pcreg\}$ we have $\npair[n]{(\Phi_S.\reg(r),\Phi_T.\reg(r))} \in \lrrg{\trust}(R_R(r))$.
   To this end, we need to consider each of cases
@@
 
-    \item TODO (the remaining cases)
+  \item $\Phi_S'(\pcreg) = \Phi_T'(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$:
+    Follows by definition of $\updPcAddr{}$ and using the fact that $r_i \neq \pcreg$ for all $i$ and the corresponding assumption of thi slemma.
+
+  \item $\npair[n-1]{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$:
+    Follows by Lemma~\ref{lem:downwards-closed} from the corresponding assumption of this lemma.
+
+  \item $\npair[n-1]{(\Phi_S'.\reg,\Phi_T'.\reg)} \in \lrrg{\trust}(W_R)$:
+    See above.
+
+  \item $\memSat[n-1]{\Phi_S'.\mem,\Phi_S'.\stk,\Phi_S'.\ms_\stk,\Phi_T'.\mem}{W_M}$:
+    These components are all unchanged from $\Phi_S$ and $\Phi_T$, so the result follows by Lemma~\ref{lem:downwards-closed} from the corresponding assumption of this lemma.
+
+  \item $W_\pcreg \oplus W_R \oplus W_M$ is defined:
+    Follows from the corresponding assumption of this lemma.
+
+  \item Theorem~\ref{thm:ftlr} holds for all $n' < n-1$:
+    follows from the corresponding assumption of this lemma.
   \end{itemize}

### 2018-06-30 fully-abstract-comp/ftlr-proof-cases.tex
Discharge another TODO.

@@
 \begin{proof}
-  \lau{TODO}
+  From $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$,
+  we get a $d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal}$ such that the following holds:
+  \begin{itemize}
+  \item $\mscode$ has no hidden calls
+  \item $\sigrets \mathrel{\#} \sigcloss$
+  \item $\sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a)$
+  \item $\forall a \in \dom(\mscode) \ldotp \sigrets,d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a$
+  \item $\exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset$
+  \end{itemize}
+
+  Particularly, we have $\sigrets,d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a$ and $\sigrets,d_\sigma(a'),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a'$.
+  From these, and the facts that $\mscode([\aaddr..\aaddr+\calllen-1]) = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}}$ and $\mscode([\aaddr'..\aaddr'+\calllen-1]) = \overline{\scall{\offpc',\offsigma'}{r_1'}{r_2'}}$, it follows that
+  \begin{itemize}
+  \item $\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ and $\sigma_\baddr+\offsigma \in d_\sigma(a)$
+  \item $\mscode(a'+\offpc') = \seal{\sigma_\baddr',\sigma_\eaddr',\sigma_\baddr'}$ and $ \sigma_\baddr'+\offsigma' \in d_\sigma(a')$
+  \end{itemize}
+  From the facts that $\sigma = \sigma_b + \offsigma = \sigma_b' + \offsigma'$, $\sigma_\baddr+\offsigma \in d_\sigma(a)$, $ \sigma_\baddr'+\offsigma' \in d_\sigma(a')$ and the disjointness of the $d_\sigma(a)$, we get that $a = a'$.
+  With this fact, the rest of the proof obligations follow directly from the other assumed equations.
 \end{proof}

### 2018-07-01 fully-abstract-comp/popl2019/paper.tex
Propose a name for our calling convention: StackTokens.

@@
 %% Title information
-\title{Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
+\title{StackTokens: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
 % \titlenote{with title note}
@@
 \begin{abstract}
-We propose and study a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
+We propose and study StackTokens: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
 The calling convention is based on linear capabilities, a type of capabilities that are prevented by the hardware from being duplicated.
@@
 
-In this work, we propose and study another way to enforce well-bracketed control flow and local state encapsulation with a single shared stack.
-Instead of CHERI's local capabilities, we build on \emph{linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
+In this work, we propose and study StackTokens: an alternative calling convention that enforces well-bracketed control flow and local state encapsulation with a single shared stack.
+Instead of CHERI's local capabilities, it builds on \emph{linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
 These capabilities are enforced by the hardware to be non-duplicable.
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities, together with some security checks, then allows us to guarantee well-bracketed control flow and local stack frame encapsulation.
+The non-duplicability of linear capabilities, together with some security checks, then allows us to guarantee well-bracketed control flow and local stack frame encapsulation, without large overhead on boundary crossings and particularly no need for clearing large blocks of memory.
 
@@
 This independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
+
+\paragraph{Outline} Blabla
 

### 2018-07-01 fully-abstract-comp/ftlr-proof-cases.tex
Sealed case.

--- 2018-06-29 fully-abstract-comp/ftlr-proof-cases.tex
@@
       \item \label{case:ftlr:move-cca-etc:move-eq}(move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_T(r_1)$ and $k = 1$.
-      \item \label{case:ftlr:move-cca-etc:cseal}(cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' =
-        \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) =
-        \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k=1$
+      \item \label{case:ftlr:move-cca-etc:cseal}(cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' = \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \le \sigma \le \sigma_\eaddr$, and $k=1$, and either $\Phi_S$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ or $\ta \# [\baddr,\eaddr]$.
       \item \label{case:ftlr:move-cca-etc:split-normal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
@@
 
+  \lau{TODO: look through this case and find the place where the ''no seal cross boundary thing'' should have been used.
+    (I have done this, and it does not seem necessary (?))}
   In case \ref{case:ftlr:call}, we may assume $r_1 \neq r_2$\footnote{If the register contains a data capability, then the execution fails in the step after the jump.
@@
 
-  Case \ref{case:ftlr:scall:reg}:
-
-  First, from reasonability of $\Phi_S$, we get that $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n-1$ steps for all $r \neq \pcreg$.
-  Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} then tells us that
-  $\npair[n-1]{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrr(W_R)$.
-
-  We then need to split the ownership of $W_R'$.
-  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
+  Case \ref{case:ftlr:scall:reg}: We need to split the ownership of $W_R'$.
+  From assumption $\npair[n]{(\Phi_S.reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we get a way to split the ownership of $W_R$.
   We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership (i.e.\ $\purePart{W_{R,1}'}$).
@@
 
-  we split the world in the same way for the registers that remain unchanged, and we get from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity} that
-  $\npair[n-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrrg{\trust}(\{\rdata,\rstk,\rtmp{1},r_1,r_2,\rretd,\rretc\})(W_R'')$ for the appropriate $W_R''$.
-
-  To obtain $\npair[n-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$, it remains to prove the following cases:
+  we split the world in the same way for the registers that remain unchanged, the
+  result follows from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity}.
+  \dominique{No: we need to use Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} together with the reasonability of registers at a boundary crossing (if there is one).}
+
+  This leaves the following cases:
   \begin{description}
@@
 
-    TODO
+    Pick $R_R' = R_R$.
+    For $r \neq r_1,r_2$, we have
+    \[
+      \npair[n-1]{(\Phi_S(r),\Phi_T(r))} \in \lrv(R_R(r))
+    \]
+    by assumption and Lemma~\ref{lem:downwards-closed}.
+
+    For $r_1$ use Lemma~\ref{lem:lrv-relates-linearity} and consider the following 2 cases:
+    \begin{itemize}
+    \item $\ta \# [\baddr,\eaddr]$: by assumption this entails $\trust = \untrusted$.
+
+      By assumption we have $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\untrusted}(R_R(r_2))$ which gives us the following facts:
+      \begin{itemize}
+      \item $[\sigma_\baddr,\sigma_\eaddr] \# (\gsigrets \uplus \gsigcloss)$
+      \item $\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \pwheap[R_R(r_2)] \ldotp \pwheap[R_R(r_2)](r) = (\pure,\_,H_\sigma) \wedge H_\sigma \; \sigma' \nequal (\lrvg{\untrusted} \circ \xi)$
+      \end{itemize}
+
+      This means that for $\sigma$ there is a region $r$ for which $\pwheap[R_R(r_2)](r) = (\pure,\_,H_\sigma)$ and $H_\sigma \; \sigma' \nequal (\lrvg{\untrusted} \circ \xi)$.
+      Pick this as the region in the sealed case and $\sigrets' = \emptyset$,
+      $\sigcloss = [\sigma_\baddr,\sigma_\eaddr]$, and $\mscode = [\baddr,\eaddr] \mapsto 0$.
+      We now need to show the following:
+      \begin{itemize}
+      \item $H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma$
+
+        To this end let $\hat{W}$ be given and show
+        \[
+          H_\sigma \; \sigma \; \hat{W} \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \; \hat{W}
+        \]
+        By transitivity of $n$-equality it suffices to show $H_\sigma \; \sigma' \; \hat{W} \nequal \lrvg{\untrusted} \circ \xi(\hat{W})$, which follows by assumption and $H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \; \hat{W} \nequal \lrvg{\untrusted} \circ \xi(\hat{W})$ which follows by definition of $H^\mathrm{code,\square}_\sigma$ and the fact that $\sigma \in \sigcloss$ and $\ta \# [\baddr,\eaddr]$.
+
+      \item $\npair[n']{(\Phi_S(r_1),\Phi_T(r_1))} \in H_\sigma \; \sigma \; \xi^{-1}(R_R(r_1)) \text{ for all $n' < n$}$
+
+        which corresponds to showing $\npair[n']{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\untrusted}(R_R(r_1))$ which is true by assumption and Lemma~\ref{lem:downwards-closed}.
+      \item $(\isLinear{\vsc_S} \Rightarrow \forall W' \future R_R(r_1), W_o, n' < n-1, \npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \npair[n']{\vsc_S,\vsc_S',\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))$
+
+        Let $W' \future R_R(r_1)$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o)$.
+        Further let $n'' \le n'$ be given and assume $\npair[n'']{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$, $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
+
+        Now consider the following cases:
+        \begin{itemize}
+        \item $\exec{\vsc_S'}$:
+
+          In this case, pick $\Phi_S' = \Phi_T' = \failed$ (as xjump fails).
+          It is trivial to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \item $\nonExec{\vsc_S'}$:
+
+          In this case consider what $\Phi_S(r_1)$ is:\\
+          If $\Phi_S(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx, \rx\}$:
+
+          If $\perm = \rwx$, then we have a contradiction with $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\untrusted}(R_R(r_1))$.
+
+          If $\perm = \rx$, then by Lemma~\ref{lem:safe-cap-exec-is-normal} we have a contradiction with $\isLinear{\Phi_S(r_1)}$.
+          \\
+          \\
+          Otherwise (not $\Phi_S(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx, \rx\}$):\\
+
+          In this case, pick $\Phi_S' = (\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk)$ and $\Phi_T' = (\ms_T,\reg_T\update{\pcreg}{\Phi_T(r_1)}\update{\rdata}{\vsc_T'})$.
+          In this case, the next step of execution fails which makes it trivial
+          to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \end{itemize}
+      \item $(\nonLinear{\src{\vsc_S}} \Rightarrow \forall W' \future \purePart{R_R(r_2)}, W_o, n' < n-1, \npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \npair[n']{\vsc_S,\vsc_S',\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))$
+
+        Let $W' \future \purePart{R_R(r_1)}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o)$.
+        Further let $n'' \le n'$ be given and assume $\npair[n'']{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$, $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
+
+        Now consider the following cases:
+        \begin{itemize}
+        \item $\exec{\vsc_S'}$:
+
+          In this case, pick $\Phi_S' = \Phi_T' = \failed$ (as xjump fails).
+          It is trivial to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \item $\nonExec{\vsc_S'}$:
+
+          In this case consider what $\Phi_S(r_1)$ is.\\
+          If $\Phi_S(r_1) = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$ and $\perm' \in \{\rwx, \rx\}$:
+
+          If $\perm' = \rwx$, then we have a contradiction with $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\untrusted}(R_R(r_1))$.
+
+          If $\perm' = \rx$, then the result follows from \ref{lem:untrusted-sealed-capability}.
+        \end{itemize}
+      \end{itemize}
+    \item $\ta \subseteq [\baddr,\eaddr]$: by assumption this entails $\trust = \trusted$.
+
+      Further, we know $\Phi_S$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$, so by the reasonability assumption on $\Phi_S$, we know $\sigma \in \gsigcloss$ and one of the following holds:
+      \begin{itemize}
+      \item $\exec{\Phi(r_1)}$ and $\Phi(r_1)$ behaves reasonably up to $n - 1$ steps.
+      \item $\nonExec{\Phi(r_1)}$ and $\Phi(r_1)$ is reasonable up to $n - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$
+      \end{itemize}
+      By $\sigma \in \gsigcloss$ we can conclude that $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \not\in \lrvg{\untrusted}(W)$ which means that the assumption $\npair{(\Phi_S(r_2),\Phi_T(r_2))} \in \lrvg{\trusted}(W)$ gives us $r \in \dom(\pwheap)$ such that $\pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\mscode,\gc}$, $[\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets \cup \sigcloss)$ and $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$, $\dom(\mscode) \subseteq \ta$.
+
+      Now pick $r$ and show
+      \begin{itemize}
+      \item $\npair[n']{(\Phi_S(r_1),\Phi_T(r_1))} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(R_R(r_1)) \text{ for all $n' < n-1$}$
+
+        This amounts to showing (1) assuming $\exec{\Phi_S(r_1)}$ show $\npair[n']{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trusted}(R_R(r_1))$, which is true by assumption and Lemma~\ref{lem:downwards-closed}.
+
+        (2) Assuming $\nonExec{\Phi_S(r_1)}$ show $\npair[n']{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\untrusted}(R_R(r_1))$, which follows by assumption, Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} and Lemma~\ref{lem:downwards-closed}.
+      \item
+        \[
+          \begin{array}{l}
+            (\isLinear{\Phi_S(r_1)} \Rightarrow \\
+            \quad\forall W' \future R_R(r_1), W_o, n' < n, \npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+            \qquad \npair[n']{\Phi_S(r_1),\vsc_S',\Phi_T(r_1),\vsc_T'} \in \lrexj(W'\oplus W_o))
+          \end{array}
+        \]
+
+        Let $W' \future R_R(r_1)$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o)$.
+        Further let $n'' \le n'$ be given and assume $\npair[n'']{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$, $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
+
+        Now consider the following cases:
+        \begin{itemize}
+        \item $\exec{\vsc_S'}$:
+
+          In this case, pick $\Phi_S' = \Phi_T' = \failed$ (as xjump fails).
+          It is trivial to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \item $\nonExec{\vsc_S'}$:
+
+          In this case consider what $\Phi_S(r_1)$ is:\\
+          If $\Phi_S(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx, \rx\}$:
+
+          If $\perm = \rwx$, then we have a contradiction with $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trusted}(R_R(r_1))$.
+
+          If $\perm = \rx$, then by Lemma~\ref{lem:safe-cap-exec-is-normal} we have a contradiction with $\isLinear{\Phi_S(r_1)}$.
+          \\
+          \\
+          Otherwise (not $\Phi_S(r_1) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx, \rx\}$):\\
+
+          In this case, pick $\Phi_S' = (\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk)$ and $\Phi_T' = (\ms_T,\reg_T\update{\pcreg}{\Phi_T(r_1)}\update{\rdata}{\vsc_T'})$.
+          In this case, the next step of execution fails which makes it trivial to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \end{itemize}
+      \item
+        \[
+          \begin{array}{l}
+            (\nonLinear{\Phi_S(r_1)} \Rightarrow \\
+            \quad \forall W' \future \purePart{R_R(r_1)}, W_o, n' < n, \npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+            \qquad \npair[n']{\Phi_S(r_1),\vsc_S',\Phi_T(r_1),\vsc_T'} \in \lrexj(W'\oplus W_o))
+          \end{array}
+        \]
+
+        Let $W' \future \purePart{R_R(r_1)}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o)$.
+        Further let $n'' \le n'$ be given and assume $\npair[n'']{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$, $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
+
+        Now consider the following cases:
+        \begin{itemize}
+        \item $\exec{\vsc_S'}$:
+
+          In this case, pick $\Phi_S' = \Phi_T' = \failed$ (as xjump fails).
+          It is trivial to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \item $\nonExec{\vsc_S'}$:
+
+          In this case consider what $\Phi_S(r_1)$ is.\\
+          If $\Phi_S(r_1) = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$ and $\perm' \in \{\rwx, \rx\}$:
+
+          If $\perm' = \rwx$, then we have a contradiction with $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\trusted}(R_R(r_1))$.
+
+          If $\perm' = \rx$, then in the case where $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \in \lrvg{\untrusted}(R_R(r_1))$, the result follows from \ref{lem:untrusted-sealed-capability}.
+
+          In the case where $\npair{(\Phi_S(r_1),\Phi_T(r_1))} \not\in \lrvg{\untrusted}(R_R(r_1))$, we want to show $(\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk)$ behaves reasonable up to $n-1$ steps.
+
+          By assumption we have $\Phi_S(r_1)$ behaves reasonably up to $n-1$ steps, so it SFTS $\reg_S(r)$ is reasonable up to $n-1$ steps in memory $\ms_S$ and free stack $\ms_\stk$ for $r \neq \pcreg$ which follows from Lemma~\ref{lem:untrusted-source-values-are-reasonable}.
+
+          By assumption we have $\npair{[\baddr',\eaddr']} \in \xReadCond{R_R(r_1)}$, so using Lemma~\ref{lem:non-expansiveness} and Lemma~\ref{lem:monotonicity} as well as $[\baddr',\eaddr'] \subseteq \ta$.
+          Using this with the other assumptions and the IH, we get:
+          \[
+            \npair[n-1]{((\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk),(\ms_T,\reg_T\update{\pcreg}{\Phi_T(r_1)}\update{\rdata}{\vsc_T'}))} \in \lro
+          \]
+
+          Otherwise (not $\Phi_S(r_1) = ((\perm',\lin'),\baddr',\eaddr',\aaddr')$ and $\perm' \in \{\rwx, \rx\}$):
+
+          In this case, pick $\Phi_S' = (\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk)$ and $\Phi_T' = (\ms_T,\reg_T\update{\pcreg}{\Phi_T(r_1)}\update{\rdata}{\vsc_T'})$.
+          In this case, the next step of execution fails which makes it trivial to show $\npair{\Phi_S',\Phi_T'} \in \lro$.
+        \end{itemize}
+      \end{itemize}
+    \end{itemize}
   \item Case \ref{case:ftlr:move-cca-etc:split-normal}:
@@
 
-    From $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we know that $\npair{(w,w')} \in \lrvg{\trust}(W_{R,2})$, $\npair{(w_2,w_2')} \in \lrvg{\trust}(\purePart{W_{R,2}]})$ and $\npair{(\Phi_S.\reg\update{r_2}{w_2},\Phi_T.\reg\update{r_2}{w_2'})} \in \lrrg{\trust}(W_R')$ with $W_R =W_R' \oplus W_{R,2}$ (using Lemma~\ref{lem:purePart-oplus} and~\ref{lem:purePart-duplicable}).
-
-    From reasonability of $\Phi_S$, we know that $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n-1$ steps.
-    Lemma~\ref{lem:trusted-and-reasonable-is-untrusted} then tells us that $\npair{(w,w')} \in \lrv(W_{R,2})$.
-
-    The result then follows from Lemma~\ref{lem:store-reg-works}, redistributing ownership in the obvious way.
+    Follows from Lemma~\ref{lem:store-reg-works}, redistributing ownership in the obvious way.
 
@@
 
+\begin{lemma}
+  \label{lem:safe-cap-exec-is-normal}
+  If
+  \begin{itemize}
+  \item $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+  \item $\perm = \{\rx,\rwx\}$
+  \item $\npair{(c,\_)} \in \lrv{\trust}(W)$
+  \end{itemize}
+  then
+  \[
+    \lin = \normal
+  \]
+
+\end{lemma}
+\begin{proof}
+  Follows from the definition.
+\end{proof}
+
+\begin{lemma}
+  \label{lem:untrusted-sealed-capability}
+  If
+  \begin{itemize}
+  \item $c_c = ((\rx,\normal),\baddr,\eaddr,\aaddr)$
+  \item $\aaddr \in [\baddr,\eaddr]$
+  \item $\npair{[\baddr,\eaddr]} \in \execCond{W}$
+  \item $\nonExec{c_d}$
+  \item $\npair{(c_d,c_d')} \in \lrvg{\untrusted}(W_o)$
+  \item $\npair{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$
+  \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
+  \item $\Phi_S = (\ms_S,\reg_S\update{\pcreg}{c_c}\update{\rdata}{c_d},\stk,\ms_\stk)$
+  \item $\Phi_T = (\ms_T,\reg_T\update{\pcreg}{c_c'}\update{\rdata}{c_d'})$
+  \item $W' \future \purePart{W}$
+  \item $W' \oplus W_o \oplus W_R \oplus W_M$ is defined
+  \end{itemize}
+  then
+  \[
+    \npair[n-1]{(\Phi_S,\Phi_T)} \in \lro
+  \]
+\end{lemma}
+\begin{proof}
+  From Lemma~\ref{lem:monotonicity} and Lemma~\ref{lem:non-linear-pure} we get
+  \[
+    \npair{[\baddr,\eaddr]} \in \execCond{W}
+  \]
+  from which we get
+  \[
+    \npair[n-1]{(c_c,c_c')} \in \lre(W')
+  \]
+  From $\npair{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$, $\npair{(c_d,c_d')} \in \lrvg{\untrusted}(W_o)$, and $W' \oplus W_o \oplus W_R \oplus W_M$ is defined, we conclude
+  \[
+    \npair{(\reg_S\update{\rdata}{c_d},\reg_T\update{\rdata}{c_d})} \in \lrrg{\untrusted}(W_R \oplus W_o)
+  \]
+  along with $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$ and Lemma~\ref{lem:non-expansiveness} we can now conclude
+  \[
+    \npair[n-1]{(\Phi_S,\Phi_T)} \in \lro
+  \]
+\end{proof}
+
 %%% Local Variables:

### 2018-07-01 fully-abstract-comp/technical-report.tex
Updated seal instruction to actually check against seal range.
Do not forget to use reasonability in the store case...
Sealed case.

@@
                                       \Phi(r_2) = \seal{\sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
-                                      \vsc = \sealed{\sigma,\Phi(r_2)} \tand\\
-                                      w_2 = \linCons{\Phi(r_2)}                                                           \end{array}
+                                      \sigma_\baddr \le \sigma \le \sigma_\eaddr \tand \\
+                                      \vsc = \sealed{\sigma,\Phi(r_1)}
+                                    \end{array}
                                     \\
@@
       \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \tand \\
+      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
       \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
@@
     \begin{itemize}
-    \item $\src{\Phi}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
+    \item $\src{\Phi}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
     \end{itemize}
@@
     \item $\src{\Phi}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$
+
+      \lau{What is this used for?}
     \item $\sigma \in \gsigcloss$ and one of the following holds:
       \begin{itemize}
-      \item $\exec{\Phi(r_2)}$ and $n' = 0$ or $\Phi(r_2)$ behaves reasonably up to $n' - 1$ steps.
-      \item $\nonExec{\Phi(r_2)}$ and $n' = 0$ or $\Phi(r_2)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$.
+      \item $\exec{\Phi(r_1)}$ and $n' = 0$ or $\Phi(r_1)$ behaves reasonably up to $n' - 1$ steps.
+      \item $\nonExec{\Phi(r_1)}$ and $n' = 0$ or $\Phi(r_1)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$.
       \end{itemize}

### 2018-07-01 fully-abstract-comp/popl2019/paper.tex
Some work on explaining the remaining problems with stack and return capabilities.

@@
 %                         %% http://ctan.org/pkg/booktabs
-% \usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-%                         %% http://ctan.org/pkg/subcaption
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
+                        %% http://ctan.org/pkg/subcaption
+\usepackage{wrapfig}
+
+\input{preamble-illu}
 
@@
 
-Well-bracketed control flow expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
+Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
 At the assembly level, this is not so obvious, as invoked functions get direct access to return pointers, that they are supposed to jump to a single time, at the end of their execution, but there is no guarantee that untrusted assembly code respects this intended usage.
@@
 
-Local state encapsulation is the guarantee that when a function invokes another function, its local variables will not have been modified when the invoked function returns.
+Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not have been read or modified when the invoked function returns.
 At the assembly level, this property is also far from obvious.
 The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
-However, untrusted assembly code is free to ignore this requirement and overwrite the local state of its caller or other stack frames.
+However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
 
@@
 
-To enforce well-bracketed control flow and local state encapsulation on a capability machine, there are essentially two approaches.
+To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
 A first approach uses separate stacks for distrusting components, and a central, trusted stack manager component that mediates cross-component invocations.
@@
 A more scalable approach retains a single stack that is shared between components.
-Enforce well-bracketed control flow and stack frame encapsulation in this approach, requires a way to temporarily provide stack and return capabilities to an untrusted component and revoke them after it returns.
+Enforcing WBCF and LSE in this approach, requires a way to temporarily provide stack and return capabilities to an untrusted component and revoke them after it returns.
 While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
 For example, CHERI offers a form of \emph{local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
-In previous work, we have demonstrated that by making the stack and return pointer local, and a number of security checks and measures, well-bracketedness and local stack frame encapsulation can be guaranteed~\citep{skorstengaard_reasoning_2017}.
+In previous work, we have demonstrated that by making the stack and return pointer local, and a number of security checks and measures, the two properties can be guaranteed~\citep{skorstengaard_reasoning_2017}.
 However, a problem with this approach is that revoking the local stack and return capabilities on every boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
 
-In this work, we propose and study StackTokens: an alternative calling convention that enforces well-bracketed control flow and local state encapsulation with a single shared stack.
+In this work, we propose and study StackTokens: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
 Instead of CHERI's local capabilities, it builds on \emph{linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
@@
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities, together with some security checks, then allows us to guarantee well-bracketed control flow and local stack frame encapsulation, without large overhead on boundary crossings and particularly no need for clearing large blocks of memory.
+The non-duplicability of linear capabilities, together with some security checks, then allows us to guarantee WBCF and LSE, without large overhead on boundary crossings and particularly no need for clearing large blocks of memory.
 
@@
 To accomodate this, we define a second operational semantics for our target language, which has a native well-bracketed stack and primitive ways to do calls and returns.
-This well-behaved semantics natively guarantees well-bracketed control flow and local stack frame encapsulation for components using our calling convention.
+This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
 As such, these components can be sure that they will only ever interact with well-behaved other components that respect our desirable properties.
@@
 \section{Linear Stack and Return Capabilities}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.49\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
+      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
+
+      % Stack pointer 1
+      \begin{scope}
+        \clip (4.6,4) rectangle (9,13);
+        \capbracebot{(4.6,4)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 1}
+      \end{scope}
+      Stack pointer 2
+      \begin{scope}
+        \clip (4.8,4) rectangle (9,13.5);
+        \capbrace{(5.2,9)}{(5.2,14)}{adv. stack\\\footnotesize cap. 2}
+      \end{scope}
+
+      \draw[red,thick,->] (4.5,10) to[out=0,in=50] node[midway,right] {} (7.5,7.3);
+      \draw[red,thick,->] (7,7.3) to[out=115,in=0] node[midway,right] {} (4.5,8);
+    \end{tikzpicture}
+    \caption{An adversary uses a previous stack frame's stack pointer.}
+    \label{fig:stack-ptr-abuse}
+  \end{subfigure}
+  \begin{subfigure}{0.49\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
+      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
+      \fill[red,draw=red,opacity=.4] (-.1,3.9) rectangle (4.6,9.1);
+      \draw[red,thick,->] (4.5,10.5) to[out=0,in=0] node[midway,right,align=center] {return\\ to wrong\\ return ptr} (4.5,3);
+    \end{tikzpicture}
+    \caption{An adversary jumps to a previous stack frame's stack pointer.}
+    \label{fig:ret-ptr-abuse}
+  \end{subfigure}
+
+  \caption{Possible ways to abuse stack and return capabilities}
+  \label{fig:stack-ret-ptr-abuse}
+\end{figure}
+
+Our StackTokens calling convention is based on a traditional single stack, shared between all components.
+As we are on a capability machine, it is natural to add some extra protection to stack and return pointers.
+First, we replace stack pointers with stack capabilities.
+When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
+Return pointers, on the other hand, are replaced by a sealed pair of return capabilities.
+They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
+
+While the above adds extra protection, it is important to understand that the above is not sufficient for enforcing WBCF and LSE.
+While adversaries only get access to capabilities that they are allowed to use during the stack frame, they have ways to store these capabilities and use them when they are no longer allowed to do so.
+Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
+The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) blabla
+
 \begin{itemize}
-\item informally explain how an adversary may try to abuse stack and return pointers
+\item informally explain how an adversary may try to abuse stack and return caps
 \item informally explain how we prevent this using linear capabilities

### 2018-07-01 fully-abstract-comp/popl2019/preamble-illu.tex
Some work on explaining the remaining problems with stack and return capabilities.

--- 2018-05-08 toplas/preamble-illu.tex
@@
     \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5,align=center] {\footnotesize #4};
     \endonlyenv
@@
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5,align=center] {\footnotesize #4};
   }{
@@
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5,align=center] {\footnotesize #4};
     \endonlyenv
@@
   \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw[fill=white] #2 rectangle #3 node[pos=.5,align=center] {\footnotesize #4};
   }{
-   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5] {#4};
+   \draw<#1>[fill=white] #2 rectangle #3 node[pos=.5,align=center] {#4};
  }
@@
     \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw<#1> #2 rectangle #3 node[pos=.5,align=center] {\footnotesize #4};
   }
@@
 
-\newcommand{\capbrace}[3][sp1]{
+\newcommand{\capbrace}[4][sp1]{
   \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-  #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
-
+  #2 -- #3 node (#1) [black,midway,xshift=1cm,align=center] {\footnotesize #4};
 }
 
-\newcommand{\stdstackstart}{
+\newcommand{\capbracebot}[4][sp1]{
+  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  #2 -- #3 node (#1) [black,near start,xshift=1cm,align=center] {\footnotesize #4};
+}
+
+\newcommand{\capbracebotred}[4][sp1]{
+  \draw [decorate,red,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  #2 -- #3 node (#1) [red,near start,xshift=1cm,align=center] {\footnotesize #4};
+}
+
+\newcommand{\stdstackstart}[1][11]{
   \scope
-    \clip (-.1,-.1) rectangle (4.6,11.1);
-    \fill[fill=white] (0,0) rectangle (4.5,11);
-    \draw (0,0) -- (0,11);
-    \draw (4.5,0) -- (4.5,11);
+    \clip (-.1,-.1) rectangle (4.6,#1 + .1);
+    \fill[fill=white] (0,0) rectangle (4.5,#1);
+    \draw (0,0) -- (0,#1);
+    \draw (4.5,0) -- (4.5,#1);
     \draw[fill=gray!50] (0,-.5) rectangle (4.5,2)
@@
   \draw (-1.5,0) node {};
-  \draw (-1.5,15) node {};
+  \draw (-1.5,#1 + 1) node {};
 }

### 2018-07-02 fully-abstract-comp/technical-report.tex
Inserted recursive domain equation.

@@
 \label{subsec:worlds}
-\[
-  \World = \Worldh \times \Worlds \times \Worldfs
-\]
-with $\Worlds$, $\Worldh$, and $\Region$ defined as follows
+\begin{theorem}
+  \label{thm:recursive-domain-eq}
+  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ and there exists an isomorphism $\xi$ such that
+  \[
+    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
+  \]
+  and for $\hat{W},\hat{W}' \in \Wor$
+  \[
+    \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})
+  \]
+\end{theorem}
+
+Where $\Worlds$, $\Worldh$, and $\Worldfs$ are defined as follows
 \[
@@
   \Regionh = 
-  \{\pure \} \times (\World \monnefun \URel{\MemSeg^2}) \times \\
-  (\Seal \parfun \World \monnefun \URel{\SealableCaps \times \SealableCaps})
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
 \end{multline*}
@@
   \begin{array}{l}
-    \{\spatial \} \times (\World \monnefun \URel{\MemSeg^2}) \uplus \\
-    \{\spatialo \} \times (\World \monnefun \URel{\MemSeg^2})\uplus \\ 
+    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
     \{\revoked\}
@@
     \end{itemize}
-    Then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ step
+    Then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ steps.
   \item \emph{Don't leak private stuff...} If
@@
   \begin{itemize}
-  \item $\npair{(w,\_)} \in \lrvg{\trusted}(W_w)$
+  \item $\npair{(w,w')} \in \lrvg{\trusted}(W_w)$
   \item $w$ is reasonable up to $n$ steps in memory $\ms_S$ and free stack $\ms_\stk$, with respect to $\code,\ta,\gsigrets,\gsigcloss$.
@@
   \begin{itemize}
-  \item $\npair{(w,\_)} \in \lrv(W_w)$
+  \item $\npair{(w,w')} \in \lrv(W_w)$
   \end{itemize}

### 2018-07-02 fully-abstract-comp/popl2019/paper.tex
Anonymized a sentence.

@@
 For example, CHERI offers a form of \emph{local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
-In previous work, we have demonstrated that by making the stack and return pointer local, and a number of security checks and measures, the two properties can be guaranteed~\citep{skorstengaard_reasoning_2017}.
+\citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and a number of security checks and measures, the two properties can be guaranteed.
 However, a problem with this approach is that revoking the local stack and return capabilities on every boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.

### 2018-07-02 fully-abstract-comp/technical-report.tex
Comment out outdated section with component layout and linking.

@@
 
-\subsection{Component layout and linking}
-\dominique{21-3-2018: TODO: update this section (further) to reflect recent changes.}
-A target- or source-level component needs a number of things.
-In the following, we list what it needs followed by a description of how they are made available.
-
-A component needs access to the following:
-\begin{description}
-\item[Code] The instructions of a component should of course be available so they can be executed.
-\item[Component seals] Each call/return point of the component, as well as every location where new closures are created, needs a unique seal in order to enforce well-bracketedness.
-  This seal needs to be unique on the entire machine.
-\item[Stack] A component stores the value of local variables on the stack. 
-\item[Persistent data storage] A component may also need to use "global variables" these are stored in a piece of memory separate from the stack.
-\item[Access to other components' exported functions] A component may need to call other components.
-  Every component makes available a number of exported values (e.g.
+% \subsection{Component layout and linking}
+% \dominique{21-3-2018: TODO: update this section (further) to reflect recent changes.}
+% A target- or source-level component needs a number of things.
+% In the following, we list what it needs followed by a description of how they are made available.
+
+% A component needs access to the following:
+% \begin{description}
+% \item[Code] The instructions of a component should of course be available so they can be executed.
+% \item[Component seals] Each call/return point of the component, as well as every location where new closures are created, needs a unique seal in order to enforce well-bracketedness.
+%   This seal needs to be unique on the entire machine.
+% \item[Stack] A component stores the value of local variables on the stack. 
+% \item[Persistent data storage] A component may also need to use "global variables" these are stored in a piece of memory separate from the stack.
+% \item[Access to other components' exported functions] A component may need to call other components.
+%   Every component makes available a number of exported values (e.g.
 sealed code-data pairs for invoking certain functions) and components are allowed to include these exported values in their data memory.
-  They could, for example, keep a linking table there through which they can invoke other components' exports.
-\end{description}
-
-In the following, we describe the setup in detail.
-Figure~\ref{fig:trg-prog-link} illustrates the setup (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).
-
-The \emph{code} and \emph{program seals} are available through a normal read-execute capability in the $\pc$ register.
-The program seals are stored as part of the code memory and can be accessed through the $\pc$ register.
-The data memory may contain references to references exported by other components, organised, for example, as a linking table.
-The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
-
-The \emph{linking table} contains pairs of sealed code and data capabilities.
-So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
-
-The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.}
-The seals are globally unique in the sense that no other program has access to this range of seals.
-
-The \emph{stack} is available through a capability in the register $r_\stk$.
+%   They could, for example, keep a linking table there through which they can invoke other components' exports.
+% \end{description}
+
+% In the following, we describe the setup in detail.
+% Figure~\ref{fig:trg-prog-link} illustrates the setup (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).
+
+% The \emph{code} and \emph{program seals} are available through a normal read-execute capability in the $\pc$ register.
+% The program seals are stored as part of the code memory and can be accessed through the $\pc$ register.
+% The data memory may contain references to references exported by other components, organised, for example, as a linking table.
+% The remaining addresses governed by the capability in the $\pc$ register contain the instructions of the program.
+
+% The \emph{linking table} contains pairs of sealed code and data capabilities.
+% So the first two addresses of the linking table contains the first pair, the two next contains the next pair and so on.
+
+% The \emph{program seals} contain at least a seal for each of the ``call/return points'' in the program.\lau{it is not clear what this means for a target level program that does not contain the call instruction.}
+% The seals are globally unique in the sense that no other program has access to this range of seals.
+
+% The \emph{stack} is available through a capability in the register $r_\stk$.
 The stack pointer is a linear read write capability.
@@
 
-The \emph{persistent data storage} is expected to be accessible through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
+% The \emph{persistent data storage} is expected to be accessible through a normal read-write capability in $r_\data$ during normal execution (during a return the stack capability will initially be in the $r_\data$ register, but it should be spliced with the returned stack and moved to the $r_\stk$ register).
 There are two ways for a program to get this data capability: (1) the data part of the code/data pair (for either a callback or a global function in the linking table) should be the sealed data capability.
 (2) During a call, the data capability should be stored on the stack, so it can be retrieved afterwards.
-\begin{figure}
-  \centering
-  \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
-  \caption{Linking and program layout (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).}
-  \label{fig:trg-prog-link}
-\end{figure}
+% \begin{figure}
+%   \centering
+%   \includegraphics[angle=90,width=\textwidth]{img/linking.pdf}
+%   \caption{Linking and program layout (modulo the fact that the code memory contains a reference to the linking table, which is now no longer allowed).}
+%   \label{fig:trg-prog-link}
+% \end{figure}
 

### 2018-07-02 fully-abstract-comp/technical-report.tex
Comment out outdated Notes section.

@@
 
-\section{Notes}
-\subsection{Leuven stay conclusions}
-\begin{enumerate}
-\item Enter capabilities replaced by sealed code/data pairs \label{item:first-point}
-  \begin{itemize}
-  \item To allow us to forbid dynamic code generation
-  \end{itemize}
-\item Conditional full-abstraction
-  \begin{itemize}
-  \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction.
+% \section{Notes}
+% \subsection{Leuven stay conclusions}
+% \begin{enumerate}
+% \item Enter capabilities replaced by sealed code/data pairs \label{item:first-point}
+%   \begin{itemize}
+%   \item To allow us to forbid dynamic code generation
+%   \end{itemize}
+% \item Conditional full-abstraction
+%   \begin{itemize}
+%   \item No undefined behaviour by \emph{trusted code} in any context implies full-abstraction.
 (Blame idea: undefined as undef of current pc, so current pc can be checked for whether it was the trusted code).
-  \item Avoids dynamic checks to protect trusted code against itself.
-  \item Possible due to point \ref{item:first-point}.
-  \end{itemize}
-\item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
-  \begin{itemize}
-  \item Allows back translation to be embedding into source language.
-    \begin{itemize}
-    \item Interpreter for back translation is not able to accurately replicate code length.
-    \end{itemize}
-  \end{itemize}
-\item Worlds similar to CSF paper, but invariants on seals as well.
-  \begin{itemize}
-  \item Invariants says what can be sealed with a certain seals.
-  \end{itemize}
-\item Linker: resolves symbols
-  \begin{itemize}
-  \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
-  \item Import refs
-  \item Fresh seal requirement
-  \end{itemize}
-\end{enumerate}
-
-\subsection{Some explanation}
-\subsubsection{The source language}
-The whole idea of this paper is to show that we enforce well-bracketed control flow correctly by taking a source language with a native stack and native call and return instructions and well-bracketedness enforced by the operational semantics.
-We then prove that we can compile this source language to a regular assembly target language (without a native stack) in a fully abstract way using our stack and return pointer discipline.
-So the design goals for our source language were precisely these: it should be an intermediate language that stays as close as possible to our target language, can be fully abstractly compiled to the target and feature a native well-behaved stack.
-This source language can then be used as the second-to-last language in the compilation chain of a fully abstract compiler.
-By compiling in a fully abstract way to our source language and then composing with our compiler, fully abstract compilation to our target language follows easily.
-
-These design goals drove us to the current definition of our source language and some special design choices.
-Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
-The reason is that we are working in a language where the length of code blocks can be observable.
-Imagine two trusted blocks of code that looks as follows:
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-p1:       <assert that r1 contains an executable capability to a block of 
-           instr_length (push) + instr_length(jmp)> 
-          <load return capability to p1-ret1 in r2>
-          <load value 0 in r3>
-          <jump to r1>
-p1-ret1:  <assert that 0 is loaded in the next free stack position>
-          <repeat the above with value 1 in r3>
-          ...
-p1-ret2:  <assert that 1 is loaded in the next free stack position>
-          <repeat the above with value 2 in r3>
-          ...
-p1-ret3:  <assert that 2 is loaded in the next free stack position>
-          halt
-
-p2:       <assert that r1 contains an executable capability to a block of 
-           instr_length (push) + instr_length(jmp)>
-          <load return capability to p2-ret1 in r2>
-          <load value 1 in r3>
-          <jump to r1>
-p2-ret1:  <assert that 1 is loaded in the next free stack position>
-          <repeat the above with value 2 in r3>
-          ...
-p2-ret2:  <assert that 2 is loaded in the next free stack position>
-          halt
-\end{lstlisting}
-
-p1 and p2 each accept an (unprotected) closure from the context whose length is restricted.
-Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
-\begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
-closure: push r3
-         jmp r2
-\end{lstlisting}
-
-However, this is no longer true in the target language.
-Assume that $\mathtt{push r3}$ is implemented as a store to $\rstk$ followed by an increment of $\rstk$.
-The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
-\begin{lstlisting}[basicstyle=\targetcolor{}\ttfamily] 
-evil-closure: jnz r3 r2
-              store $\rstk$ r3
-              jmp r2
-\end{lstlisting}
-If we now assume that the instructions in evil-closure (jnz+store+jmp) are the same length as the original closure(store+increment+jmp), then the adversary can now construct a closure that discriminates $\mathtt{p1}$ and $\mathtt{p2}$ above.
-The essential point here is that an assumption in the source language that relies on facts about instruction length (``a closure restricted to the instruction-size of push+ret that modifies the stack and returns can only be implemented in one way'') is no longer valid, because there are now more low-level ways of interacting with the stack.
-
-To remedy this, we decided to modify our source language to use a different way of interacting with the stack.
-This alternative approach stays even closer to the target language.
-The problem is not actually that we shouldn't offer push/pop and call/return instructions, but rather that our language should also offer an equivalent for the more low-level interactions with the stack that are possible in the target language.
-As such, we introduce special token values that play the role of the stack pointer/return pointer and we allow all interactions with them that are possible using instructions like \sstore{}{}, \smove{}{} etc.
-Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the sealed part of a code and data return closure.
-Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:syntax}, this has the semantic effect that one would expect.
-Similarly, rather than invoking \sreturn{}, one would jump to the $\retptrd$ and $\retptrc$ sealed code/data pair.
-
-We do not offer primitive \sreturn{}, \spush{} and \spop{} instructions, even though there is nothing keeping us from doing that, either as an additional primitive instruction or as a macro on top of the primitive token values mentioned above.
-We do offer a primitive \scall{\offpc,\offsigma}{}{} instruction, and it offers the only way to create a new stack frame.
-It is the only instruction that does not have a target language counterpart.
-However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{\offpc,\offsigma}{}{} compiles to.
-We will discuss later that this means that if we back translate an adversary piece of code (in the full abstraction proof), then it will be back translated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
-In other words, the back-translated code never uses call to protect its own stack frame from the code that it passes control to.
-This means that the back-translated code is quite insecure, but we don't need it to be: it just needs to be behaviorally equivalent to the target language code that it represents.
-
-% Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
-% \lau{07-09-2017: One would use $\scall{\offpc,\offsigma}{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
+%   \item Avoids dynamic checks to protect trusted code against itself.
+%   \item Possible due to point \ref{item:first-point}.
+%   \end{itemize}
+% \item replace push/pull/call/ret/sload by symbolic return pointer (pair) and stack pointer.
+%   \begin{itemize}
+%   \item Allows back translation to be embedding into source language.
+%     \begin{itemize}
+%     \item Interpreter for back translation is not able to accurately replicate code length.
+%     \end{itemize}
+%   \end{itemize}
+% \item Worlds similar to CSF paper, but invariants on seals as well.
+%   \begin{itemize}
+%   \item Invariants says what can be sealed with a certain seals.
+%   \end{itemize}
+% \item Linker: resolves symbols
+%   \begin{itemize}
+%   \item Export refs (Question: do we allow other things than sealed code/data pairs to be exported?)
+%   \item Import refs
+%   \item Fresh seal requirement
+%   \end{itemize}
+% \end{enumerate}
+
+% \subsection{Some explanation}
+% \subsubsection{The source language}
+% The whole idea of this paper is to show that we enforce well-bracketed control flow correctly by taking a source language with a native stack and native call and return instructions and well-bracketedness enforced by the operational semantics.
+% We then prove that we can compile this source language to a regular assembly target language (without a native stack) in a fully abstract way using our stack and return pointer discipline.
+% So the design goals for our source language were precisely these: it should be an intermediate language that stays as close as possible to our target language, can be fully abstractly compiled to the target and feature a native well-behaved stack.
+% This source language can then be used as the second-to-last language in the compilation chain of a fully abstract compiler.
+% By compiling in a fully abstract way to our source language and then composing with our compiler, fully abstract compilation to our target language follows easily.
+
+% These design goals drove us to the current definition of our source language and some special design choices.
+% Particularly, contrary to what one might expect in a language with a native stack, we do not use call/return and push/pop instructions.
+% The reason is that we are working in a language where the length of code blocks can be observable.
+% Imagine two trusted blocks of code that looks as follows:
+% \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+% p1:       <assert that r1 contains an executable capability to a block of 
+%            instr_length (push) + instr_length(jmp)> 
+%           <load return capability to p1-ret1 in r2>
+%           <load value 0 in r3>
+%           <jump to r1>
+% p1-ret1:  <assert that 0 is loaded in the next free stack position>
+%           <repeat the above with value 1 in r3>
+%           ...
+% p1-ret2:  <assert that 1 is loaded in the next free stack position>
+%           <repeat the above with value 2 in r3>
+%           ...
+% p1-ret3:  <assert that 2 is loaded in the next free stack position>
+%           halt
+
+% p2:       <assert that r1 contains an executable capability to a block of 
+%            instr_length (push) + instr_length(jmp)>
+%           <load return capability to p2-ret1 in r2>
+%           <load value 1 in r3>
+%           <jump to r1>
+% p2-ret1:  <assert that 1 is loaded in the next free stack position>
+%           <repeat the above with value 2 in r3>
+%           ...
+% p2-ret2:  <assert that 2 is loaded in the next free stack position>
+%           halt
+% \end{lstlisting}
+
+% p1 and p2 each accept an (unprotected) closure from the context whose length is restricted.
+% Together with the checks performed by p1 and p2, there is only one possible implementation of the closure that will not fail, namely
+% \begin{lstlisting}[basicstyle=\sourcecolor{}\ttfamily] 
+% closure: push r3
+%          jmp r2
+% \end{lstlisting}
+
+% However, this is no longer true in the target language.
+% Assume that $\mathtt{push r3}$ is implemented as a store to $\rstk$ followed by an increment of $\rstk$.
+% The adversary can now construct a closure that does not use the standard implementation of push, but instead does this:
+% \begin{lstlisting}[basicstyle=\targetcolor{}\ttfamily] 
+% evil-closure: jnz r3 r2
+%               store $\rstk$ r3
+%               jmp r2
+% \end{lstlisting}
+% If we now assume that the instructions in evil-closure (jnz+store+jmp) are the same length as the original closure(store+increment+jmp), then the adversary can now construct a closure that discriminates $\mathtt{p1}$ and $\mathtt{p2}$ above.
+% The essential point here is that an assumption in the source language that relies on facts about instruction length (``a closure restricted to the instruction-size of push+ret that modifies the stack and returns can only be implemented in one way'') is no longer valid, because there are now more low-level ways of interacting with the stack.
+
+% To remedy this, we decided to modify our source language to use a different way of interacting with the stack.
+% This alternative approach stays even closer to the target language.
+% The problem is not actually that we shouldn't offer push/pop and call/return instructions, but rather that our language should also offer an equivalent for the more low-level interactions with the stack that are possible in the target language.
+% As such, we introduce special token values that play the role of the stack pointer/return pointer and we allow all interactions with them that are possible using instructions like \sstore{}{}, \smove{}{} etc.
+% Specifically, as shown in \cref{sec:domains}, there are new token capabilities $\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}$, $\retptrd$ and $\retptrc$ which represent a copy of the stack pointer, and the sealed part of a code and data return closure.
+% Rather than invoking \spush{} or \spop{}, one can store into the $\stkptr{}$ and modify its current address and as we can see in \cref{sec:syntax}, this has the semantic effect that one would expect.
+% Similarly, rather than invoking \sreturn{}, one would jump to the $\retptrd$ and $\retptrc$ sealed code/data pair.
+
+% We do not offer primitive \sreturn{}, \spush{} and \spop{} instructions, even though there is nothing keeping us from doing that, either as an additional primitive instruction or as a macro on top of the primitive token values mentioned above.
+% We do offer a primitive \scall{\offpc,\offsigma}{}{} instruction, and it offers the only way to create a new stack frame.
+% It is the only instruction that does not have a target language counterpart.
+% However, there is a counterpart to \sxjmp{}{}, the instruction that would be used for jumping in the code that \scall{\offpc,\offsigma}{}{} compiles to.
+% We will discuss later that this means that if we back translate an adversary piece of code (in the full abstraction proof), then it will be back translated (under our trivial back-translation) to equivalent code that \emph{never} does an actual call, but always just jumps to where it wants to go.
+% In other words, the back-translated code never uses call to protect its own stack frame from the code that it passes control to.
+% This means that the back-translated code is quite insecure, but we don't need it to be: it just needs to be behaviorally equivalent to the target language code that it represents.
+
+% % Finally, rather than invoking call, \dominique{7-9-2017: err, what's the reason we don't follow the same pattern for call?}.
+% % \lau{07-09-2017: One would use $\scall{\offpc,\offsigma}{}{}$ with a pair of sealed capabilities to securely jump to this pair - that is to have this pair executed in a new stack frame.
 The return would happen just by using $\sxjmp{}{}$ with the $\retptrd$ and $\retptrc$.
@@
 In other words, we need something to recognise when a return happens.}
-% \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{\offpc,\offsigma}{}{}$ instruction.
+% % \lau{07-09-2017: On second thought, when compiled all the code necessary for the call will be generated based on the $\scall{\offpc,\offsigma}{}{}$ instruction.
 The real implementation of the $\retptrc$ will simply point to a piece of clean up code that splices together the old/private stack pointer with the one we got from the untrusted party.
@@
 
-Note that the token values representing the stack and return pointer in our source language are essentially just a weird alternative syntax for something we would otherwise write using call/return and push/pop.
-However, it does remove the problem described above of relying on facts that follow from code length, because we can now write exactly the same code in source and target and it will do exactly the same thing if it is well-behaved.
-The code would look exactly the same in source and target, it would just be working with token values in the source versus real stack and return pointers in the target.
-That the problem is really gone should follow from being able to write a full-abstraction proof.
-However, our approach should also simplify that proof.
-The reason is that back-translating should become essentially trivial: rather than performing some form of emulation or a real back-translation (which would inevitably make code size larger), it should be sufficient to just use unmodified target instructions as source instructions that will just be working with different types of values.
-
-\subsubsection{Conditional full abstraction}
-\label{sec:conditional-full-abstraction}
-
-Another novelty in this work is the fact that we do not want to prove general full abstraction (i.e.
+% Note that the token values representing the stack and return pointer in our source language are essentially just a weird alternative syntax for something we would otherwise write using call/return and push/pop.
+% However, it does remove the problem described above of relying on facts that follow from code length, because we can now write exactly the same code in source and target and it will do exactly the same thing if it is well-behaved.
+% The code would look exactly the same in source and target, it would just be working with token values in the source versus real stack and return pointers in the target.
+% That the problem is really gone should follow from being able to write a full-abstraction proof.
+% However, our approach should also simplify that proof.
+% The reason is that back-translating should become essentially trivial: rather than performing some form of emulation or a real back-translation (which would inevitably make code size larger), it should be sufficient to just use unmodified target instructions as source instructions that will just be working with different types of values.
+
+% \subsubsection{Conditional full abstraction}
+% \label{sec:conditional-full-abstraction}
+
+% Another novelty in this work is the fact that we do not want to prove general full abstraction (i.e.
 two source programs are behaviorally equivalent if and only if they are compiled to behaviorally equivalent target programs).
-The reason is that we only want to prove a result about programs that respect the calling convention, i.e.\ that do not do anything stupid.
-For example, in the old setup with global/local capabilities, when trusted code performs an \scall{\offpc,\offsigma}{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
-Such sanity conditions are necessary (we cannot guarantee anything for code that does such stupid things) but we do not want to spend computation time on checking them.
-Instead, we expect a reasonable compiler to guarantee in some way that its generated code never does such stupid things and then our guarantees should apply.
-
-Note that in the new setup with linear capabilities instead of local/global ones, there seems to be less opportunity for the trusted code to do stupid things, so it may be that we no longer need these sanity conditions, but this is not clear yet.
-
-Assuming that we need sanity conditions on the source code, the question is how we formulate them in our full abstraction.
-Our idea is to formulate a notion of conditional full abstraction: ``if source programs t1 and t2 never perform stupid things under any possible source interaction, then they are source-equivalent iff their compilations are target-equivalent.''.
-However, the question remains how we formalise the condition ``if source programs t1 and t2 never perform stupid things under any possible source interaction''.
-
-Two possible approaches:
-\begin{itemize}
-\item \emph{Blame}: the idea here would be that we introduce a notion of ``undefined behavior'' in the operational semantics.
-  ``Doing stupid things'' like leaving copies of the stack pointer around on a call, would cause the system to enter an undefined state.
-  We would then like to express the above mentioned sanity condition as ``t1 and t2 never evaluate to undefined when plugged into an arbitrary program context''.
-  However, that doesn't work, because it does not discriminate between undefined behavior caused by t1 or t2 (on the one hand) and the program context (on the other hand).
-  To remedy this, we intended to use a notion of blame assignment: the undefined state would contain the address of the pc that was active at the moment it went wrong.
-  We would then express the above mentioned sanity condition as ``t1 and t2 never evaluate to undefined(pc) where pc is within the trusted code's assigned code memory when plugged into an arbitrary program context''.
-
-  The problem with this approach is that it does not combine well with dynamically generated code (solution: exclude code-generation?) and that it adds a lot of complexity.
-
-\item \emph{A logical relation}: An alternative approach (suggested by Lars to Lau) might be to define a logical relation to express the sanity condition ``t1 and t2 never evaluate to undefined when plugged into an arbitrary program context''.
-  This remains to be thought through in more detail, but it might be simpler and/or be easier to combine with dynamic code generation.
-\end{itemize}
-
-\subsection{Linear Capabilities}
-\label{sec:linear-cap}
-The idea with linear capabilities is to limit aliasing of a capability.
+% The reason is that we only want to prove a result about programs that respect the calling convention, i.e.\ that do not do anything stupid.
+% For example, in the old setup with global/local capabilities, when trusted code performs an \scall{\offpc,\offsigma}{}{}, it should make sure that it does not store a copy of its stack pointer in a place that is accessible to the callee (e.g.\ in a register or in the callee's part of the stack).
+% Such sanity conditions are necessary (we cannot guarantee anything for code that does such stupid things) but we do not want to spend computation time on checking them.
+% Instead, we expect a reasonable compiler to guarantee in some way that its generated code never does such stupid things and then our guarantees should apply.
+
+% Note that in the new setup with linear capabilities instead of local/global ones, there seems to be less opportunity for the trusted code to do stupid things, so it may be that we no longer need these sanity conditions, but this is not clear yet.
+
+% Assuming that we need sanity conditions on the source code, the question is how we formulate them in our full abstraction.
+% Our idea is to formulate a notion of conditional full abstraction: ``if source programs t1 and t2 never perform stupid things under any possible source interaction, then they are source-equivalent iff their compilations are target-equivalent.''.
+% However, the question remains how we formalise the condition ``if source programs t1 and t2 never perform stupid things under any possible source interaction''.
+
+% Two possible approaches:
+% \begin{itemize}
+% \item \emph{Blame}: the idea here would be that we introduce a notion of ``undefined behavior'' in the operational semantics.
+%   ``Doing stupid things'' like leaving copies of the stack pointer around on a call, would cause the system to enter an undefined state.
+%   We would then like to express the above mentioned sanity condition as ``t1 and t2 never evaluate to undefined when plugged into an arbitrary program context''.
+%   However, that doesn't work, because it does not discriminate between undefined behavior caused by t1 or t2 (on the one hand) and the program context (on the other hand).
+%   To remedy this, we intended to use a notion of blame assignment: the undefined state would contain the address of the pc that was active at the moment it went wrong.
+%   We would then express the above mentioned sanity condition as ``t1 and t2 never evaluate to undefined(pc) where pc is within the trusted code's assigned code memory when plugged into an arbitrary program context''.
+
+%   The problem with this approach is that it does not combine well with dynamically generated code (solution: exclude code-generation?) and that it adds a lot of complexity.
+
+% \item \emph{A logical relation}: An alternative approach (suggested by Lars to Lau) might be to define a logical relation to express the sanity condition ``t1 and t2 never evaluate to undefined when plugged into an arbitrary program context''.
+%   This remains to be thought through in more detail, but it might be simpler and/or be easier to combine with dynamic code generation.
+% \end{itemize}
+
+% \subsection{Linear Capabilities}
+% \label{sec:linear-cap}
+% The idea with linear capabilities is to limit aliasing of a capability.
 We specifically wanted a linear capability for the stack pointer, to not have to erase large pieces of memory, which we have to when we use local capabilities.
@@
 
-There are multiple ways that linear capabilities could work (we choose the third one):
-\begin{enumerate}
-\item A linear capability comes with a system wide guarantee that no \emph{linear} alias exists for this capability (or any part of the memory it has authority over).
+% There are multiple ways that linear capabilities could work (we choose the third one):
+% \begin{enumerate}
+% \item A linear capability comes with a system wide guarantee that no \emph{linear} alias exists for this capability (or any part of the memory it has authority over).
 In other words, the linear capability is the only \emph{linear} capability for this piece of memory.
@@
 
-In order for this to work, the memory must initially be partitioned into two parts.
+% In order for this to work, the memory must initially be partitioned into two parts.
 One controlled by a non-linear capability and one controlled by a linear capability.
@@
 
-In this scheme, it would be okay to permit linear capabilities to become non-linear (which would consume the linear capability).
+% In this scheme, it would be okay to permit linear capabilities to become non-linear (which would consume the linear capability).
 The reasoning would be that it is okay to remove the aliasing restriction and thus make the part of memory with the no-aliasing guarantee smaller.
@@
 
-\item A linear capability can be created from a non-linear capability.
+% \item A linear capability can be created from a non-linear capability.
 The linear capability has no guarantee that it is not aliased, but no aliases can be made from the linear capability.
@@
 
-In this scheme, the system would initially have one non-linear capability for all of the memory. The linear capabilities would then be generated as they are needed.
-\item A linear capability is ``born'' with the guarantee that it is not aliased, either by linear or non-linear capabilities.
+% In this scheme, the system would initially have one non-linear capability for all of the memory. The linear capabilities would then be generated as they are needed.
+% \item A linear capability is ``born'' with the guarantee that it is not aliased, either by linear or non-linear capabilities.
 This guarantee is maintained by the semantics.
@@
 
-Note that in this scheme, it would be good if malloc provided a way to initially obtain a linear capability in this model.
-Otherwise, there would never be any linear capabilities other than stack pointers, which seems a waste.
-\end{enumerate}
-Notice that the two first schemes are incompatible as they would undermine each other.
-
-We chose to use the second scheme for the following reasons:
-\begin{itemize}
-\item While the stack pointer was the motivation for the linear capabilities, we believe they can also be used for other things such as return pointers.
+% Note that in this scheme, it would be good if malloc provided a way to initially obtain a linear capability in this model.
+% Otherwise, there would never be any linear capabilities other than stack pointers, which seems a waste.
+% \end{enumerate}
+% Notice that the two first schemes are incompatible as they would undermine each other.
+
+% We chose to use the second scheme for the following reasons:
+% \begin{itemize}
+% \item While the stack pointer was the motivation for the linear capabilities, we believe they can also be used for other things such as return pointers.
 When used for return pointers, we know that the return pointer used to return to us was the only one available.
@@
 This would be undesirable if we want to provide several entities access to the same program.
-\end{itemize}
-We later abandoned the second option for the following reason:
-\begin{itemize}
-\item The initial idea with linear capabilities was to use them for stack pointers, because there would only ever be one capability for a any part of the stack.
+% \end{itemize}
+% We later abandoned the second option for the following reason:
+% \begin{itemize}
+% \item The initial idea with linear capabilities was to use them for stack pointers, because there would only ever be one capability for a any part of the stack.
 However, when normal capabilities can become linear, then it is only the creator of the linear capability that can be sure that no alias is created from the linear capability.
@@
 
-In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from an untrusted source.
-\end{itemize}
-
-Even though we call them linear, they are more like \emph{affine} capabilities because our current operational semantics allows for them to be overwritten.
+% In other words, we needed a system-wide guarantee to be able to put some trust in linear capabilities from an untrusted source.
+% \end{itemize}
+
+% Even though we call them linear, they are more like \emph{affine} capabilities because our current operational semantics allows for them to be overwritten.
 We choose to stick with the name as the term linear is also being misused by others for instance in Rust.
 
-\subsubsection{Write-linear capabilities}
-
-At the time is writing, it is unclear whether write-linear (in the style of write-local permission\footnote{Here it would mean that linear capabilities can only be written through capabilities with write-linear permission and all capabilities with write-linear permission must themselves be linear.}) is necessary.
+% \subsubsection{Write-linear capabilities}
+
+% At the time is writing, it is unclear whether write-linear (in the style of write-local permission\footnote{Here it would mean that linear capabilities can only be written through capabilities with write-linear permission and all capabilities with write-linear permission must themselves be linear.}) is necessary.
 One might think that it is necessary to prevent ``indirect aliasing'', which here would mean that a linear capability would be stored in a piece of memory for which there are multiple non-linear capabilities.
@@
 
-One thing write-linear capabilities may be useful for is to make sure that return capabilities are used in the correct order, i.e., ensuring well-bracketedness. 
-\lau{08-09-2017: I have not thought the above completely through} 
-\lau{13-09-2017: Dominique wrote something about this in an e-mail.
+% One thing write-linear capabilities may be useful for is to make sure that return capabilities are used in the correct order, i.e., ensuring well-bracketedness. 
+% \lau{08-09-2017: I have not thought the above completely through} 
+% \lau{13-09-2017: Dominique wrote something about this in an e-mail.
 The splicing necessary for each call would be what is needed to make the write-linear obsolete.
@@
 
-We have also considered removing local capabilities as these no longer seem necessary and only add complexity.
+% We have also considered removing local capabilities as these no longer seem necessary and only add complexity.
 However, they may be relevant to ensure the well-bracketed control flow (this will need further consideration).
 
-\dominique{14-9-2017: Currently, we have removed local capabilities and we believe write-linear will (a) not work (because it would require memory clearing like write-local) and (b) not be needed (because of the splicing discussed above).}
-
-\subsection{Opaque capabilities}
-At some point, we contemplated having opaque capabilities.
+% \dominique{14-9-2017: Currently, we have removed local capabilities and we believe write-linear will (a) not work (because it would require memory clearing like write-local) and (b) not be needed (because of the splicing discussed above).}
+
+% \subsection{Opaque capabilities}
+% At some point, we contemplated having opaque capabilities.
 Opaque capabilities would hide the addresses of the capability, i.e, a getter on the base, end or address would yield nothing.
@@
 These could be:
-\begin{enumerate}
-\item reset current address of capability to its base address
-\item return difference between current address and base address of a capability
-\item increase/decrease current address of a capability by a set amount
-\end{enumerate}
-(Linear capabilities may still call for some of these to be added for practical purposes. (like when the executing capability is linear, then it cannot be moved from the pc register).)
-
-At the moment of writing this, I am not sure why we introduced them, but it was related to hiding program lengths (and stack height).
-At this juncture, it does, however, not seem necessary to have opaque because sealed capabilities can do their job.
-
-\subsection{Calling convention design decisions}
-\subsubsection{Returning the full stack}
-When a callee return from a call, they must return all of the stack they were passed.
+% \begin{enumerate}
+% \item reset current address of capability to its base address
+% \item return difference between current address and base address of a capability
+% \item increase/decrease current address of a capability by a set amount
+% \end{enumerate}
+% (Linear capabilities may still call for some of these to be added for practical purposes. (like when the executing capability is linear, then it cannot be moved from the pc register).)
+
+% At the moment of writing this, I am not sure why we introduced them, but it was related to hiding program lengths (and stack height).
+% At this juncture, it does, however, not seem necessary to have opaque because sealed capabilities can do their job.
+
+% \subsection{Calling convention design decisions}
+% \subsubsection{Returning the full stack}
+% When a callee return from a call, they must return all of the stack they were passed.
 If we omit this requirement, then we cannot guarantee well-bracketedness.
@@
 
-\begin{itemize}
-\item An adversary calls our trusted code with a call-back. 
-\item Our code uses part of the stack and calls the callback with the rest of the stack.
-\item The adversary splits that stack in two and saves the part of the stack adjacent to our stack in some persistent memory.
+% \begin{itemize}
+% \item An adversary calls our trusted code with a call-back. 
+% \item Our code uses part of the stack and calls the callback with the rest of the stack.
+% \item The adversary splits that stack in two and saves the part of the stack adjacent to our stack in some persistent memory.
 The adversary calls us anew with a callback and the part of the stack they did not save.
-\item We use part of the stack we receive and call the callback with the rest of the stack.
+% \item We use part of the stack we receive and call the callback with the rest of the stack.
 The adversary can now use the saved stack to return from the first call to the callback breaking well-bracketedness.
 The reason this is possible is because we do not check the size of the stack.
-\end{itemize}
-Figure~\ref{fig:ret-full-stk} illustrates the above example.
+% \end{itemize}
+% Figure~\ref{fig:ret-full-stk} illustrates the above example.
 In the figure, $f$ is the function to ``is'' and $g$ is the callback passed to ``us''.
@@
 The stack that is used to return lines up with $\var{Us}_\priv$, so the return is successful.
-\begin{figure}
-  \centering
-  \includegraphics[angle=90,width=\textwidth]{img/ret-full-stk.pdf}
-  \caption{Illustration of the stack in the example that illustrates why the entire stack must be returned.}
-  \label{fig:ret-full-stk}
-\end{figure}
-
-\subsubsection{Restriction on stack allocation}
-We need to somehow make sure that it is always the same stack that is used.
+% \begin{figure}
+%   \centering
+%   \includegraphics[angle=90,width=\textwidth]{img/ret-full-stk.pdf}
+%   \caption{Illustration of the stack in the example that illustrates why the entire stack must be returned.}
+%   \label{fig:ret-full-stk}
+% \end{figure}
+
+% \subsubsection{Restriction on stack allocation}
+% We need to somehow make sure that it is always the same stack that is used.
 If we don't, then an adversary can simply split the stack in two, use one part for one call and the other for another call.
@@
 
-\begin{itemize}
-  \item An adversary starts the execution. They split the stack in two and call us with one part (say the top part) along with a callback.
-  \item We use part of the stack and call the adversary with the rest of the stack.
-  \item The adversary calls us again this time using the other part of the stack (here the bottom part of the stack).
-  \item Again, we use part of the stack and call the adversary with the rest of this part of the stack.
-\end{itemize}
-At this point, the adversary can return from either of the two calls. Swapping around the order in which the adversary uses the two parts of the stack changes nothing.
-
-The example is illustrated in Figure~\ref{fig:stk-alloc}.
-
-One way to solve this problem is to make the "top address" of the stack known. There are many ways to do this, but we have chosen to do the following:
-\begin{itemize}
-\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
-\item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
-\item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
-\item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
-\end{itemize}
-
-\begin{figure}
-  \centering
-  \includegraphics[angle=270,trim={0 0 5cm 0},width=\textwidth]{img/stk-alloc.pdf}
-  \caption{Illustration of the potential stack allocation issue.}
-  \label{fig:stk-alloc}
-\end{figure}
+% \begin{itemize}
+%   \item An adversary starts the execution. They split the stack in two and call us with one part (say the top part) along with a callback.
+%   \item We use part of the stack and call the adversary with the rest of the stack.
+%   \item The adversary calls us again this time using the other part of the stack (here the bottom part of the stack).
+%   \item Again, we use part of the stack and call the adversary with the rest of this part of the stack.
+% \end{itemize}
+% At this point, the adversary can return from either of the two calls. Swapping around the order in which the adversary uses the two parts of the stack changes nothing.
+
+% The example is illustrated in Figure~\ref{fig:stk-alloc}.
+
+% One way to solve this problem is to make the "top address" of the stack known. There are many ways to do this, but we have chosen to do the following:
+% \begin{itemize}
+% \item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
+% \item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
+% \item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
+% \item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
+% \end{itemize}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[angle=270,trim={0 0 5cm 0},width=\textwidth]{img/stk-alloc.pdf}
+%   \caption{Illustration of the potential stack allocation issue.}
+%   \label{fig:stk-alloc}
+% \end{figure}
 

### 2018-07-02 fully-abstract-comp/popl2019/paper.tex
Reorder sentence parts for better reading flow.
Restructure paper outline

@@
   We propose and study StackTokens: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
-The calling convention is based on linear capabilities, a type of capabilities that are prevented by the hardware from being duplicated.
+  The calling convention is based on linear capabilities, a type of capabilities that are prevented from being duplicated by the hardware.
   In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, using what we call a fully abstract overlay semantics.
@@
 \item present the full abstraction theorem.
-\item sketch high-level structure of the proof
 \end{itemize}
+
+\section{Proving full abstraction}
+\label{sec:fa-proof}
+
+\begin{itemize}
+\item Logical relation
+\item FTLR
+\item Sketch high-level structure of the proof
+\end{itemize}
+
 \section{Discussion}

### 2018-07-02 fully-abstract-comp/technical-report.tex

@@
 
-\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
-  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
-}
-\dominique{TODO: better name than ``reasonable'' for all of this.}
+% \dominique{TODO: better name than ``reasonable'' for all of this.}
 Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.

### 2018-07-02 fully-abstract-comp/technical-report.tex
removed TODOs and added short disclaimer.

@@
 \tableofcontents
+Disclaimer: While the proofs in this technical report are done, the text can be lacklustre and from time to time out of date.
+We will make the technical report up to date as soon as possible.
+
 \section{The two capability machines}
@@
 \]
-\lau{TODO: Make this condition require a set of adjacent regions. Check whether necessary for the stack region.}
 where $\iota_A$ is a standard region defined in Section~\ref{sec:standard-regions}.
@@
 
-\lau{19-03-2018: Add the link capability to the above? code in ta?}
-\lau{TODO: look at the $[b,e] \subseteq ...$ condition - will it work for splice?}
 \begin{multline*}
@@
 \label{sec:full-abstraction}
-\lau{TODO: Did we agree that memory satisfaction should require $\stkb$ to be a free stack address (or at least not in the heap part)? I think we did.}
-
 \begin{definition}[Source language contextual equivalence]

### 2018-07-02 fully-abstract-comp/technical-report.tex
Uncomment up to date sections in the notes section of the TR.

--- 2018-07-02 fully-abstract-comp/technical-report.tex
@@
       \reg_S(\pcreg) \neq \retptrc(\_) \wedge 
-      \reg_S(\rdata) \neq \retptrd(\_) \wedge 
+      \reg_S(\rdata) \neq \retptrd(\_) \wedge \\
       \nonExec{\reg_S(\rdata)} \wedge
@@
 
-% \section{Notes}
+\section{Notes}
 % \subsection{Leuven stay conclusions}
@@
 
-% At the moment of writing this, I am not sure why we introduced them, but it was related to hiding program lengths (and stack height).
+% At the moment of writing this, I am not sure why we thought about them, but it was related to hiding program lengths (and stack height).
 % At this juncture, it does, however, not seem necessary to have opaque because sealed capabilities can do their job.
 
-% \subsection{Calling convention design decisions}
-% \subsubsection{Returning the full stack}
-% When a callee return from a call, they must return all of the stack they were passed.
+\subsection{Calling convention design decisions}
+\subsubsection{Returning the full stack}
+When a callee return from a call, they must return all of the stack they were passed.
 If we omit this requirement, then we cannot guarantee well-bracketedness.
@@
 
-% \begin{itemize}
-% \item An adversary calls our trusted code with a call-back. 
-% \item Our code uses part of the stack and calls the callback with the rest of the stack.
-% \item The adversary splits that stack in two and saves the part of the stack adjacent to our stack in some persistent memory.
+\begin{itemize}
+\item An adversary calls our trusted code with a call-back. 
+\item Our code uses part of the stack and calls the callback with the rest of the stack.
+\item The adversary splits that stack in two and saves the part of the stack adjacent to our stack in some persistent memory.
 The adversary calls us anew with a callback and the part of the stack they did not save.
-% \item We use part of the stack we receive and call the callback with the rest of the stack.
+\item We use part of the stack we receive and call the callback with the rest of the stack.
 The adversary can now use the saved stack to return from the first call to the callback breaking well-bracketedness.
 The reason this is possible is because we do not check the size of the stack.
-% \end{itemize}
-% Figure~\ref{fig:ret-full-stk} illustrates the above example.
+\end{itemize}
+Figure~\ref{fig:ret-full-stk} illustrates the above example.
 In the figure, $f$ is the function to ``is'' and $g$ is the callback passed to ``us''.
@@
 The stack that is used to return lines up with $\var{Us}_\priv$, so the return is successful.
-% \begin{figure}
-%   \centering
-%   \includegraphics[angle=90,width=\textwidth]{img/ret-full-stk.pdf}
-%   \caption{Illustration of the stack in the example that illustrates why the entire stack must be returned.}
-%   \label{fig:ret-full-stk}
-% \end{figure}
-
-% \subsubsection{Restriction on stack allocation}
-% We need to somehow make sure that it is always the same stack that is used.
+\begin{figure}
+  \centering
+  \includegraphics[angle=90,width=\textwidth]{img/ret-full-stk.pdf}
+  \caption{Illustration of the stack in the example that illustrates why the entire stack must be returned.}
+  \label{fig:ret-full-stk}
+\end{figure}
+
+\subsubsection{Restriction on stack allocation}
+We need to somehow make sure that it is always the same stack that is used.
 If we don't, then an adversary can simply split the stack in two, use one part for one call and the other for another call.
@@
 
-% \begin{itemize}
-%   \item An adversary starts the execution. They split the stack in two and call us with one part (say the top part) along with a callback.
-%   \item We use part of the stack and call the adversary with the rest of the stack.
-%   \item The adversary calls us again this time using the other part of the stack (here the bottom part of the stack).
-%   \item Again, we use part of the stack and call the adversary with the rest of this part of the stack.
-% \end{itemize}
-% At this point, the adversary can return from either of the two calls. Swapping around the order in which the adversary uses the two parts of the stack changes nothing.
-
-% The example is illustrated in Figure~\ref{fig:stk-alloc}.
-
-% One way to solve this problem is to make the "top address" of the stack known. There are many ways to do this, but we have chosen to do the following:
-% \begin{itemize}
-% \item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
-% \item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
-% \item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
-% \item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
-% \end{itemize}
-
-% \begin{figure}
-%   \centering
-%   \includegraphics[angle=270,trim={0 0 5cm 0},width=\textwidth]{img/stk-alloc.pdf}
-%   \caption{Illustration of the potential stack allocation issue.}
-%   \label{fig:stk-alloc}
-% \end{figure}
+\begin{itemize}
+  \item An adversary starts the execution. They split the stack in two and call us with one part (say the top part) along with a callback.
+  \item We use part of the stack and call the adversary with the rest of the stack.
+  \item The adversary calls us again this time using the other part of the stack (here the bottom part of the stack).
+  \item Again, we use part of the stack and call the adversary with the rest of this part of the stack.
+\end{itemize}
+At this point, the adversary can return from either of the two calls. Swapping around the order in which the adversary uses the two parts of the stack changes nothing.
+
+The example is illustrated in Figure~\ref{fig:stk-alloc}.
+
+One way to solve this problem is to make the "top address" of the stack known. There are many ways to do this, but we have chosen to do the following:
+\begin{itemize}
+\item The stack grows downwards, so the ``last address'' of the stack is the base address of the initial stack capability.
+\item The base address of the stack capability is a fixed address, so in the semantics, it will be expressed as a constant that is publicly known.
+\item At some point before a call, it must be checked whether the stack we are using actually has the globally known base address (if not we must fail because we cannot trust this stack).
+\item To ensure that the check is made, we require it in the semantic condition (at this moment of time not defined, so we are yet to see what it looks like).
+\end{itemize}
+
+\begin{figure}
+  \centering
+  \includegraphics[angle=270,trim={0 0 5cm 0},width=\textwidth]{img/stk-alloc.pdf}
+  \caption{Illustration of the potential stack allocation issue.}
+  \label{fig:stk-alloc}
+\end{figure}
 

### 2018-07-02 fully-abstract-comp/popl2019/paper.tex
comment
typo.
Minor updates to paper outline.
More precise
Added some things to the preamble, input preamble of tech report
Added todo about itemized contributions
Added a phantom subfigure to get some spacing
Added some text explaining the two scenarios where normal capabilities not sufficient.

@@
 
+\usepackage{todonotes}
+
 \input{preamble-illu}
+
+\input{../preamble}
 
@@
 \affiliation{
-  \institution{Aarhus University}
+  \institution{Aarhus~University}
   \country{Denmark}                    %% \country is recommended
@@
 
+\todo[inline]{Add itemized contributions}
+
 \paragraph{Outline} Blabla
@@
   \centering
-  \begin{subfigure}{0.49\linewidth}
+  \begin{subfigure}{0.4\linewidth}
     \centering
@@
   \end{subfigure}
-  \begin{subfigure}{0.49\linewidth}
+  \begin{subfigure}{0.18\linewidth}
+    \phantom{testtestes}
+  \end{subfigure}
+  \begin{subfigure}{0.4\linewidth}
     \centering
@@
 When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
-Return pointers, on the other hand, are replaced by a sealed pair of return capabilities.
+Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
 They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
 
+%informally explain how an adversary may try to abuse stack and return caps
 While the above adds extra protection, it is important to understand that the above is not sufficient for enforcing WBCF and LSE.
@@
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
-The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) blabla
+In both examples some trusted and adversarial code has been taking turns calling each other, so the stack now alternates between stack frame of the trusted and adversarial code.
+The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how the trusted code tries to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
+However, restricting the stack capability does not help when the trusted code in the first call gave access to the part of the stack where the stack frame of the second call now resides.
+Generally speaking, the trusted code has no reason to ever trust a stack capability it receives from adversarial code as the adversarial code may have duplicated the stack capability and stored it for later use.
+In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), the trusted code has given the adversarial code two sealed pairs of return capabilities, one in each of the calls to the adversarial code.
+The adversarial code may have stored the pair of sealed return pointers from the first call, so that it can access it in the second call.
+The figure illustrates how the adversarial code uses the return pair from the first call breaking WBCF.
 
@@
 \begin{itemize}
-\item present our target language and its operational semantics (excerpts)
+\item present our \emph{target language} and its operational semantics (excerpts)
 \item mention roughly what components look like
@@
 \begin{itemize}
-\item mention that tail calls are supported through xcall
+\item mention that tail calls are supported through xjmp
 \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.

### 2018-07-02 fully-abstract-comp/popl2019/preamble-illu.tex
Update layout of capability braces - add command for linear cap brace.

@@
 \newcommand{\capbrace}[4][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+  \draw [decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt},yshift=0pt]
+  #2 -- #3 node (#1) [black,midway,xshift=1cm,align=center] {\footnotesize #4};
+}
+
+\newcommand{\lincapbrace}[4][sp1]{
+  \draw [densely dotted,decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt},yshift=0pt]
   #2 -- #3 node (#1) [black,midway,xshift=1cm,align=center] {\footnotesize #4};
@@
 \newcommand{\capbracebot}[4][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  \draw [decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
   #2 -- #3 node (#1) [black,near start,xshift=1cm,align=center] {\footnotesize #4};
@@
 \newcommand{\capbracebotred}[4][sp1]{
-  \draw [decorate,red,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  \draw [decorate,red,decoration={brace,amplitude=7pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
   #2 -- #3 node (#1) [red,near start,xshift=1cm,align=center] {\footnotesize #4};

### 2018-07-02 fully-abstract-comp/popl2019/paper.tex
Update layout of capability braces - add command for linear cap brace.
Text about linear capabilities and how they mitigate the first illustration issue.

@@
       \begin{scope}
-        \clip (4.8,4) rectangle (9,13.5);
-        \capbrace{(5.2,9)}{(5.2,14)}{adv. stack\\\footnotesize cap. 2}
+        \clip (4.8,4) rectangle (9,13);
+        \capbrace{(5.2,9)}{(5.2,13.5)}{adv. stack\\\footnotesize cap. 2}
       \end{scope}
@@
 
+% informally explain how we prevent this using linear capabilities
+\lau[inline]{Write an introductory paragraph to the linear capability bit}
+
+% What is linear capabilities, split/splice
+A linear capability is a capability that cannot be duplicated.
+This is enforced dynamically, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
+A capability machine with linear capabilities also needs special instructions to split and splice.
+Split and splice do exactly what their names indicate: Split splits a capability into two capabilities with the same permissions and with ranges that are disjointed but sum up to the range of the original capability, and splice does the inverse.
+
+% Stack capability linear
+The linear capabilities are put to use by using a linear capability for the stack pointer.
+When a trusted piece of code makes a call, it splits the stack capability in two, so it has a stack capability for its stack frame and one for the unused part of the stack.
+The stack capability for the caller stack frame is sealed and used as the data part of the sealed return pair.
+The stack capability for the remainder of the stack is given to the callee.
+% Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
+As illustrated in Figure~\ref{fig:stack-ptr-abuse-prev}, this prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
+The trusted code knows that the stack capability is linear, so when it splits the stack capability, so it has a linear capability for its local stack frame, then it also knows that this part of memory cannot have an alias.
+The trusted code encapsulates the capability for its stack frame by sealing it which provides the guarantee that it can only be accessed when used with the appropriate code return pointer.
+
+% Prevention of attack 2
+
+% known stack base
+% + more
+% non-empty stack frames
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.4\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
+      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
+
+      \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
+
+      % Stack pointer 1
+      \lincapbrace{(4.6,4)}{(4.6,7)}{adv. stack\\\footnotesize cap. 1}
+      % return cap
+      \lincapbrace{(4.6,7)}{(4.6,9)}{data return\\\footnotesize cap. 2}
+     %  Stack pointer 2
+      \begin{scope}
+        \clip (4.8,4) rectangle (9,13);
+        \lincapbrace{(4.6,9)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 2}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{The non-duplicable linear stack capability for the trusted code's
+      stack frame and the opacity of sealed capabilities ensures LSE.}
+    \label{fig:stack-ptr-abuse-prev}
+  \end{subfigure}
+  \begin{subfigure}{0.18\linewidth}
+    \phantom{testtestes}
+  \end{subfigure}
+  \begin{subfigure}{0.4\linewidth}
+
+  \end{subfigure}
+\end{figure}
+
 \begin{itemize}
@@
 \end{itemize}
+
 \section{Related Work}
@@
 \item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
-\item undefined behavior and full abstraction \citep{juglaret_beyond_2016}
+\item undefined behavior and full abstraction \citep{juglaret_beyond_2016} \dominique{TODO}
+\item CheriBSD etc.
 \end{itemize}

### 2018-07-02 fully-abstract-comp/popl2019/paper.tex
Added todo about figures
More text and a figure to illustrate point in the text.

@@
 \section{Linear Stack and Return Capabilities}
-
+\todo[inline]{The stack grows downwards, maybe we should make the figures reflect this.}
 \begin{figure}
@@
 % Prevention of attack 2
+In a traditional calling convention with a single stack, the stack is used as a call stack in the sense that it keeps track of the order of the calls.
+If everyone follows the calling convention, then a call pushes a stack frame to the call stack and a return pops the topmost stack frame from the call stack.
+This convention is not followed by the adversarial code in Figure~\ref{fig:ret-ptr-abuse} as the adversarial code skips two stack frames.
+In the presence of adversarial code, we need some enforcement mechanism that forces the adversarial code to follow the call stack order.
+One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
+The callee would have to give up this token on return to prevent it from spending the same token multiple times As it turns out, the stack capability for the unused part of the stack can be seen as such a token: When the callee returns, we expect that it returns the stack capability we gave it in the call.
+If this stack capability is not present, then we have no whether our stack frame is on top of the stack, so we should refuse to continue execution.
+If we receive a stack pointer, then we try to splice it with the stack pointer for our local stack frame (which by now has been unsealed as we resume execution).
+The splice instruction only splices adjacent capabilities, so in the case where the stack frame that is returned to is on top of the stack, the returned stack capability and the unsealed stack capability are adjacent, so the splice is successful.
+In the case where the stack frame is not on top of the call stack, as illustrated in Figure~\ref{fig:ret-ptr-abuse}, part of the stack capability will be bound in the sealed return pairs for the higher stack frames making the splice fail as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}.
+
+% Where do we get the stack pointer from, how can we know it is linear
+% The other attack.
+
+% Non-empty trusted stack frames.
 
@@
   \begin{subfigure}{0.4\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
+      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
 
+      \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
+
+     %  Stack pointer 2
+      \begin{scope}
+        \clip (4.8,4) rectangle (9,13);
+        \lincapbrace{(4.6,9)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 2}
+      \end{scope}
+
+      \redcross{(5,4)}
+    \end{tikzpicture}
+    \caption{The trusted caller fails to splice the stack capability returned by
+    the adversary with the capability for the trusted caller's local stack frame.}
+    \label{fig:ret-ptr-abuse-prev}
   \end{subfigure}
+
 \end{figure}

### 2018-07-02 fully-abstract-comp/popl2019/preamble-illu.tex
More text and a figure to illustrate point in the text.

@@
 
+\newcommand{\redcross}[2][1]{
+  \draw ($#2 + (-#1/2,#1/2)$) edge[color=red,thick] ($#2 + (#1/2,-#1/2)$)
+        ($#2 + (#1/2,#1/2)$) edge[color=red,thick] ($#2 + (-#1/2,-#1/2)$);
+
+}
+
 \newcommand{\stdstackstart}[1][11]{
@@
     \draw[fill=gray!50] (0,-.5) rectangle (4.5,2)
-    node[pos=.5,color=black,align=center,yshift=0.1cm] {\footnotesize Higher
+    node[pos=.5,color=black,align=center,yshift=0.1cm] {\footnotesize Lower
       stack\\\footnotesize frames...};

### 2018-07-02 fully-abstract-comp/technical-report.tex
Fix a problem related to steps accounting in the component-related lemmas.

--- 2018-07-02 fully-abstract-comp/technical-report.tex
@@
       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
+      &\qquad\quad\text{then } \npair[n']{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \text{ for all $n' < n$} \tand\\
       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
@@
   \begin{equation*}
-    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+    \npair[n']{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
   \end{equation*}
-  and
+  for all $n' < n$ and
   \begin{equation*}
@@
   \begin{align*}
-    \npair{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1]', \mscode[1]\uplus\msdata[1]')} &\in \lrheap(\pwheap[W_1])(W')\\
-    \npair{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2]', \mscode[2]\uplus\msdata[2]')} &\in \lrheap(\pwheap[W_2])(W')
+    \npair[n']{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1]', \mscode[1]\uplus\msdata[1]')} &\in \lrheap(\pwheap[W_1])(W')\\
+    \npair[n']{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2]', \mscode[2]\uplus\msdata[2]')} &\in \lrheap(\pwheap[W_2])(W')
   \end{align*}
+  for all $n' < n$.
   It follows by Lemma~\ref{lem:combined-memory-disjoint-world} that 
   \begin{equation*}
-    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+    \npair[n']{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
   \end{equation*}
+  for all $n' < n$.
 
@@
     \begin{equation*}
-      \npair{(\sigrets\uplus\sigcloss,\mspad\uplus\mscode\uplus \msdata', \mspad\uplus\mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
+      \npair[n']{(\sigrets\uplus\sigcloss,\mspad\uplus\mscode\uplus \msdata', \mspad\uplus\mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
     \end{equation*}
+    for all $n' < n$.
+
    Take $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ such that
@@
     \begin{itemize}
-    \item $\npair{(\msdata'|_{\{\addr\}},\msdata'|_{\{\addr\}})} \in \stdreg{\{\aaddr\},\gc}{\lin}.H~ \xi^{-1}(R_W(r_\addr))$:
-      By definition, it suffices to prove that for all $n' < n$,
+    \item $\npair[n']{(\msdata'|_{\{\addr\}},\msdata'|_{\{\addr\}})} \in \stdreg{\{\aaddr\},\gc}{\lin}.H~ \xi^{-1}(R_W(r_\addr))$ for all $n' < n$:
+      Take $n' < n$.
+      By definition, it suffices to prove that,
       $\npair[n']{(\msdata'(\addr),\msdata'(\addr))} \in \lrv(R_W(r_\addr))$.
@@
       \begin{equation*}
-        \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+        \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'})
       \end{equation*}
@@
       \begin{itemize}
-      \item $\npair{A_{\mathrm{non-linear}}} \in \readCond{\normal,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
-      \item $\npair{A_{\mathrm{non-linear}}} \in \writeCond{\normal,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
-      \item $\npair{A_{\mathrm{own}}(\addr)} \in \readCond{\linear,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
-      \item $\npair{A_{\mathrm{own}}(\addr)} \in \writeCond{\linear,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair[n']{A_{\mathrm{non-linear}}} \in \readCond{\normal,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair[n']{A_{\mathrm{non-linear}}} \in \writeCond{\normal,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair[n']{A_{\mathrm{own}}(\addr)} \in \readCond{\linear,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
+      \item $\npair[n']{A_{\mathrm{own}}(\addr)} \in \writeCond{\linear,R_W(\addr)}$: follows by Lemma~\ref{lem:monotonicity} using the fact that $R_W(a) \future R_W'(a)$ and by definition, using the choice of $W$, $R_W'(\addr)$
       \item $\pwheap[R_W(\addr)](r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$: follows from the choice of $W$ and $R_W'(\addr)$ and the fact that $R_W(a) \future R_W'(a)$
@@
       \end{itemize}
-    \item $\npair{(\mspad\uplus\mscode,\mspad\uplus\mscode)} \in  \codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H~ \xi^{-1}(\purePart{W'})$:
+    \item $\npair[n']{(\mspad\uplus\mscode,\mspad\uplus\mscode)} \in  \codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H~ \xi^{-1}(\purePart{W'})$ for all $n' < n$:
+      Take $n' < n$.
       We take $\trust = \trusted$ if $[\baddr-1,\eaddr+1] \subseteq \ta$ and $\trust =\untrusted$ otherwise.
@@
         by assumption
-      \item $\forall a \in \dom(\code), n' < n \ldotp \npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{}(\xi(\xi^{-1}(\purePart{W'}))))$:
+      \item $\forall a \in \dom(\code) \ldotp \npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{}(\xi(\xi^{-1}(\purePart{W'}))))$:
         Note first that $\purePart{}(\xi(\xi^{-1}(\purePart{W'}))) =\purePart{W'}$.
@@
   \begin{equation*}
-    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
+    \npair[n']{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
   \end{equation*}
-  and
+  for all $n' < n$ and
   \begin{equation*}
@@
 
-  \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
+  \item $\memSat[n']{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$:
     Since $W_M \future W$, we have seen above that 
-    \begin{equation*}
-      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap[W_M])(W_M)
-    \end{equation*}
-    and hence by Lemma~\ref{lem:downwards-closed}:
     \begin{equation*}
@@
   If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
-  $\npair{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$.
+  $\npair[n']{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
 \end{lemma}
@@
     Take $n$ the amount of steps in the termination of $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$.
-    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\preceq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\preceq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\preceq,\gc_1](W_1\uplus W_1')$
@@
     Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \trg{\term}$.
-    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\succeq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\succeq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\succeq,\gc_2](W_2\uplus W_2')$
@@
     Take $n$ the amount of steps in the termination of $\plug{\context}{\var{comp}_1} \trg{\term}$.
-    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
@@
     Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
-    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n']{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$

### 2018-07-02 fully-abstract-comp/popl2019/paper.tex
Steal+adapt some Related Work text from the ESOP paper...

--- 2018-07-02 fully-abstract-comp/popl2019/paper.tex
@@
 
-\usepackage{todonotes}
-
 \input{preamble-illu}
-
-\input{../preamble}
 
@@
 \affiliation{
-  \institution{Aarhus~University}
+  \institution{Aarhus University}
   \country{Denmark}                    %% \country is recommended
@@
 
-\todo[inline]{Add itemized contributions}
-
 \paragraph{Outline} Blabla
@@
   \centering
-  \begin{subfigure}{0.4\linewidth}
+  \begin{subfigure}{0.49\linewidth}
     \centering
@@
       \begin{scope}
-        \clip (4.8,4) rectangle (9,13.5);
-        \capbrace{(5.2,9)}{(5.2,14)}{adv. stack\\\footnotesize cap. 2}
+        \clip (4.8,4) rectangle (9,13);
+        \capbracebot{(5.2,9)}{(5.2,13.5)}{adv. stack\\\footnotesize cap. 2}
       \end{scope}
@@
   \end{subfigure}
-  \begin{subfigure}{0.18\linewidth}
-    \phantom{testtestes}
-  \end{subfigure}
-  \begin{subfigure}{0.4\linewidth}
+  \begin{subfigure}{0.49\linewidth}
     \centering
@@
 When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
-Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
+Return pointers, on the other hand, are replaced by a sealed pair of return capabilities.
 They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
 
-%informally explain how an adversary may try to abuse stack and return caps
 While the above adds extra protection, it is important to understand that the above is not sufficient for enforcing WBCF and LSE.
@@
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
-In both examples some trusted and adversarial code has been taking turns calling each other, so the stack now alternates between stack frame of the trusted and adversarial code.
-The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how the trusted code tries to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
-However, restricting the stack capability does not help when the trusted code in the first call gave access to the part of the stack where the stack frame of the second call now resides.
-Generally speaking, the trusted code has no reason to ever trust a stack capability it receives from adversarial code as the adversarial code may have duplicated the stack capability and stored it for later use.
-In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), the trusted code has given the adversarial code two sealed pairs of return capabilities, one in each of the calls to the adversarial code.
-The adversarial code may have stored the pair of sealed return pointers from the first call, so that it can access it in the second call.
-The figure illustrates how the adversarial code uses the return pair from the first call breaking WBCF.
+The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) blabla
 
@@
 \end{itemize}
+
 \section{A Capability Machine with Sealing and Linear Capabilities}
+\label{sec:capab-mach-with}
 \begin{itemize}
-\item present our \emph{target language} and its operational semantics (excerpts)
+\item present our target language and its operational semantics (excerpts)
 \item mention roughly what components look like
 \end{itemize}
+
 \section{Formulating Security with a Fully Abstract Overlay Semantics}
+\label{sec:form-secur-with}
 \begin{itemize}
@@
 \section{Discussion}
+\label{sec:discussion}
 \begin{itemize}
@@
 \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
+\item Performance?  Implementability of linear caps?
 \end{itemize}
+
 \section{Related Work}
+
+Capability machines originate with Dennis and Van Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to Levy~\cite{Levy1984capability} and Watson et al.~\cite{Watson2015Cheri} for an overview of previous work.
+The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is modeled CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
+The latter is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
+In the interest of simplicity, we have omitted features of CHERI that were not needed for modelling StackTokens (e.g.\ local capabilities, virtual memory).
+
+Plenty of other papers enforce well-bracketed control flow at a low level, but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+This includes particularly the line of work on \emph{control-flow integrity}~\cite{abadi_control-flow_2005}.
+This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
+Contrary to StackTokens, CFI can be implemented on commodity hardware rather than capability machines.
+However, its attacker model is different and its security goals are weaker.
+They assume an attacker that is not able to execute code, but can overwrite arbitrary data at any time during execution (to model buffer overflows).
+In terms of security goals, the technique does not enforce local stack encapsulation.
+Also, it only enforces a weak form of control flow correctness, saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
+Such a property ignores temporal properties and seems hard to use for reasoning.
+
+More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}.
+Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
+TODO: adapt the following sentence...
+Also, both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, while we define a logical relation that can be used to reason about specific programs that rely on well-bracketed control flow.
+
 \begin{itemize}
 \item CFI and friends
+\item CheriBSD etc.
 \item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
-\item undefined behavior and full abstraction \citep{juglaret_beyond_2016}
+\item undefined behavior and full abstraction \citep{juglaret_beyond_2016} \dominique{TODO}
 \end{itemize}

### 2018-07-03 fully-abstract-comp/ftlr-proof-cases.tex
removed TODOs and added short disclaimer.
Fix non-expansiveness of stack, free stack and heap satisfaction, adapt step accounting throughout the paper to follow suit.

@@
 
-  The property for $\lrheap$, $\memSatFStack{\_,\_}{\_}$, $\memSatStack{\_,\_}{\_}$ follows from the downward closure of regions in the world, and the property for $(\_,\_,\_,\_) :_n W$ follows from those.
+  The property for $\lrheap$, $\memSatFStack{\_,\_}{\_}$, $\memSatStack{\_,\_}{\_}$ follows by their definition from the quantifications over $n' < n$, and the property for $(\_,\_,\_,\_) :_n W$ follows from those.
 \end{proof}
@@
   \item If $W_1 = W_1' \oplus W_1''$ then there exist $W_2' \nequal W_1'$ and $W_2'' \nequal W_1''$ such that $W_2 = W_2' \oplus W_2''$.
+  \item If $W_1' \nequal W_2'$ then $W_1\oplus W_1' \nequal W_2 \oplus W_2'$.
+  \item $\xi^{-1}(W_1) \nequal[n-1] \xi^{-1}(W_2)$
   \end{itemize}
@@
   \item $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W}$ iff $\memSat{(\ms_S,\stk,\ms_\stk,\ms_T)}{W'}$.
-  \item ...
   \end{itemize}
@@
 
-  The property for $\lrv$ follows from the other properties, by definition, from Lemma~\ref{lem:n-equality-props}, non-expansiveness of regions \dominique{Is $\xi$ non-expansive/contractive/?}
-  \lau{I believe it is at-least non-expansive and monotone as it is a morphism in the category of preordered c.o.f.e.'s.
-    Do you also ``need'' it to be contractive?}
-  TODO
+  The property for $\lrv$ follows from the other properties, by definition, from Lemma~\ref{lem:n-equality-props}, and non-expansiveness of regions.
 
@@
 
-  The property for $\lrheap$, $\memSatFStack{\_,\_}{\_}$, $\memSatStack{\_,\_}{\_}$ and $(\_,\_,\_,\_) :_n W$ follows from the non-expansiveness of regions in the world. \dominique{and $\xi?$}
-  TODO
+  The property for $\lrheap$, $\memSatFStack{\_,\_}{\_}$, $\memSatStack{\_,\_}{\_}$ and $(\_,\_,\_,\_) :_n W$ follows from the non-expansiveness of regions in the world and Lemma~\ref{lem:n-equality-props}.
 \end{proof}
@@
   $\forall r \in \activeReg{\pwheap[W_M]}$, we have that
-  $\npair{R_\ms(r)} \in  \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$.
+  $\npair[n']{R_\ms(r)} \in  \pwheap[W_\var{heap}](r).H \; \xi^{-1}(R_W(r))$ for all $n' < n$.
   We also get an $R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal}$ such that  $\biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)$.
@@
   We have that $r \in \addressable{\lin, \pwheap} \subseteq
-  \activeReg{\pwheap[W_\var{heap}]}$, so $\npair{R_\ms(r)} \in
+  \activeReg{\pwheap[W_\var{heap}]}$, so $\npair[n']{R_\ms(r)} \in
   \pwheap[W_M](r).H \; \xi^{-1}(R_W(r))$.
@@
   also $\pwheap[W_M](r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$. This
-  means that also $\npair{R_\ms(r)} \in
+  means that also $\npair[n']{R_\ms(r)} \in
   H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.
 
-  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r)$ and we get a $S : R(r) \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in R(r)} S(\aaddr)$ and $\forall \aaddr \in R(r),n' < n \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
-
-  Since $\aaddr \in R(r)$ and $n' < n$, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
+  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r)$ and we get a $S : R(r) \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in R(r)} S(\aaddr)$ and $\forall \aaddr \in R(r) \ldotp \npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
+
+  Since $\aaddr \in R(r)$, we have that $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
   we can take $W_M' = W_r' \oplus W_{\var{heap}}' \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}})$ with $W_r' = \oplus_{\aaddr \in (R(r)\setminus \{\aaddr\})} S(\aaddr)$ and $W_\var{heap}' = \oplus_{r' \in (\dom(\activeReg{\pwheap[W_\var{heap}]})\setminus \{r\})} R_W(r')$, and get
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ because $\pwheap[W_M](r)$ is address-stratified and $\pwheap[W_M](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
+  We have that $\npair[n']{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwheap[W_\var{heap}](r).H \; \xi^{-1}(W_r')$ for all $n' < n$ because $\pwheap[W_M](r)$ is address-stratified and $\pwheap[W_M](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
 
@@
   $\forall r \in \activeReg{\pwfree[W_\var{free\_stack}]}$, we have that
-  $\npair{R_\ms(r)} \in  \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
-
-  We have that $r \in \addressable{\lin, \pwfree} \subseteq \activeReg{\pwfree[W_\var{free\_stack}]}$, so $\npair{R_\ms(r)} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
+  $\npair[n']{R_\ms(r)} \in  \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$ for all $n' < n$.
+
+  We have that $r \in \addressable{\lin, \pwfree} \subseteq \activeReg{\pwfree[W_\var{free\_stack}]}$, so $\npair[n']{R_\ms(r)} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$.
   Because $\pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$ and $W\oplus W_M = W \oplus (W_\var{free\_stack} \oplus W_\var{free\_stack} \oplus W_\var{stack})$ is defined, it follows that also $\pwfree[W_\var{free\_stack}](r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H$.
-  This means that also $\npair{R_\ms(r)} \in H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.
-
-  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r)$ and we get a $S : R(r) \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in R(r)} S(\aaddr)$ and $\forall \aaddr \in R(r),n' < n \ldotp \npair[n']{(\ms_\stk(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
-
-  Since $\aaddr \in R(r)$ and $n' < n$, we have that $\npair[n']{(\ms_\stk(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
+  This means that also $\npair[n']{R_\ms(r)} \in H^{\mathrm{std}}_{R(r)}\; \xi^{-1}(R_W(r))$.
+
+  From this, it follows that $\dom(R_\ms(r).1) = \dom(R_\ms(r).2) = R(r)$ and we get a $S : R(r) \fun \World$ with $\xi(\xi^{-1}(R_W(r))) = \oplus_{\aaddr \in R(r)} S(\aaddr)$ and $\forall \aaddr \in R(r) \ldotp \npair[n']{(\ms_\stk(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$.
+
+  Since $\aaddr \in R(r)$, we have that $\npair[n']{(\ms_\stk(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))$ and
   we can take $W_M' = W_r' \oplus W_{\var{heap}} \oplus (W_{\var{stack}} \oplus W_{\var{free\_stack}}')$ with $W_r' = \oplus_{\aaddr \in (R(r)\setminus \{\aaddr\})} S(\aaddr)$ and $W_\var{free\_stack}' = \oplus_{r' \in (\dom(\activeReg{\pwheap[W_\var{heap}]})\setminus \{r\})} R_W(r')$, and get
@@
 
-  We have that $\npair{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
+  We have that $\npair[n']{(R_\ms(r).1\update{a}{0},R_\ms(r).2\update{a}{0})} \in \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(W_r')$ for all $n' < n$ because $\pwfree[W_\var{free\_stack}](r)$ is address-stratified and $\pwfree[W_\var{free\_stack}](r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$.
 
@@
           \item $\ms^{(3)}_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f}$
-          \item $\memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}''}$
-          \item $\memSatFStack{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$
-          \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}}'')$
+          \item $\memSatStack[n''']{\stk,\ms_\var{T,stack}}{W_{\var{stack}}''}$
+          \item $\memSatFStack[n''']{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$
+          \item $\npair[n''']{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}}'')$
           \end{itemize}
@@
           By assumption and the fact that the only change is that we moved part of the stack to the free stack.
-        \item $\memSatStack{(\opc_1,\ms_1):: \dots :: (\opc_m,\ms_m),\ms_\var{T,stack}\setminus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}}{W_{\var{stack}}'''}$
+        \item $\memSatStack[n''']{(\opc_1,\ms_1):: \dots :: (\opc_m,\ms_m),\ms_\var{T,stack}\setminus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}}{W_{\var{stack}}'''}$
 
@@
           All the new partitions are future worlds of the old ones as none of them owned the revoked region (it was owned by $W_R''$).
-        \item $\memSatFStack{\ms^{(3)}_\stk \uplus \ms_{\priv,S}^{(3)},\ms_\var{T,free\_stack} \uplus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}}{W_{\var{free\_stack}}'''}$
-
-          From the $\memSatFStack{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$ assumption we get $R_\ms : \dom(\activeReg{W_{\var{free\_stack}}''}) \fun \MemSeg \times \MemSeg$ and $R_W : \dom(\activeReg{W_{\var{free\_stack}}''}.\free) \fun \Worlds$ for which
+        \item $\memSatFStack[n''']{\ms^{(3)}_\stk \uplus \ms_{\priv,S}^{(3)},\ms_\var{T,free\_stack} \uplus \ms_\var{T,stack} | _{\dom(\ms_{\priv,S}^{(3)})}}{W_{\var{free\_stack}}'''}$
+
+          From the $\memSatFStack[n''']{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$ assumption we get $R_\ms : \dom(\activeReg{W_{\var{free\_stack}}''}) \fun \MemSeg \times \MemSeg$ and $R_W : \dom(\activeReg{W_{\var{free\_stack}}''}.\free) \fun \Worlds$ for which
           \begin{itemize}
@@
           \item $W_{\var{free\_stack}}'' \bigoplus_{r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free})} R_W(r)$
-          \item $\forall r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free})\ldotp \npair[n''']{R_\ms(r)} \in W_{\var{free\_stack}}''.\free(r).H \; \xi^{-1}(R_W(r))$
+          \item $\forall r \in \dom(\activeReg{W_{\var{free\_stack}}''.\free}), n'''' < n'''\ldotp \npair[n'''']{R_\ms(r)} \in W_{\var{free\_stack}}''.\free(r).H \; \xi^{-1}(R_W(r))$
           \end{itemize}
@@
           \[
-            \forall r \in \dom(\activeReg{W_{\var{free\_stack}}'''.\free}) \in \npair[n''']{R_\ms(r)} \in W_{\var{free\_stack}}'''.\free(r).H \; \xi^{-1}(R_W(r))
+            \forall r \in \dom(\activeReg{W_{\var{free\_stack}}'''.\free}),n'''' < n''' \ldotp \npair[n'''']{R_\ms(r)} \in W_{\var{free\_stack}}'''.\free(r).H \; \xi^{-1}(R_W(r))
           \]
           for $r \in \dom(\activeReg{W_{\var{free\_stack}}'''.\free}) \setminus \{r_{b_\stk'}\dots r_{e_\stk'}\}$, it follows from monotonicity of the $H$ (memory interpretation) function.
-          For $r_{a'} \in \{r_{b_\stk'}\dots r_{e_\stk'}\}$, we need to show
-          \[
-            \npair[n''']{(\ms_{\priv,S}^{(3)}(a') , \ms_\var{T,stack}|(a'))} \in \stdreg{\{a\},\gc}{\spa}.H (\xi^{-1}R_W(r_{a'}))
+          For $r_{a'} \in \{r_{b_\stk'}\dots r_{e_\stk'}\}$ and $n'''' < n'''$, we need to show
+          \[
+            \npair[n'''']{(\ms_{\priv,S}^{(3)}(a') , \ms_\var{T,stack}|(a'))} \in \stdreg{\{a\},\gc}{\spa}.H (\xi^{-1}R_W(r_{a'}))
           \]
@@
           \[
-            \npair[n''']{(\ms_{\priv,S}^{(3)}(a') , \ms_\var{T,stack}|(a'))} \in \lrv (W_{a'}''')
+            \npair[n'''']{(\ms_{\priv,S}^{(3)}(a') , \ms_\var{T,stack}|(a'))} \in \lrv (W_{a'}''')
           \]
@@
 
-        \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}}''')$
+        \item $\npair[n''']{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}}''')$
 
@@
     Follows by assumption and the fact that we have shuffled around some memory\footnote{Thanks to the register-file safety assumption we know for sure that the memory we remove from the free stack is actually there.}.
-  \item $\memSatStack{((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S,\ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}}{W_{M,\var{stack}}'}$:
+  \item $\memSatStack[n-1]{((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S,\ms_\var{T,stack} \uplus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}}{W_{M,\var{stack}}'}$:
 
@@
     \[
-      \npair[n-1]{(\ms_{\stk\_\priv,S},\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42})}
+      \npair[n']{(\ms_{\stk\_\priv,S},\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42})}
       \in
@@
     \]
+    for all $n' < n-1$.
     Where $R_W'(r_\var{priv\_stk})$ is the part of $W_{M,\var{stack}}'$ with the ownership used for addresses $[\aaddr_\stk, \eaddr_\stk]$ in the memory satisfaction assumption.
@@
     \[
-      \forall \aaddr \in [\aaddr_\stk,\eaddr_\stk] \ldotp \npair[n-1]{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
-    \]
-    for $\aaddr = \aaddr_\stk$ it is trivial as we have to show
+      \forall \aaddr \in [\aaddr_\stk,\eaddr_\stk] \ldotp \npair[n']{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
+    \]
+    for $n' < n-1$.
+    For $\aaddr = \aaddr_\stk$ it is trivial as we have to show
     \[
@@
     \[
-      \npair[n-1]{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
-    \]
+      \npair[n']{(\ms_{\stk\_\priv,S}(\aaddr),\ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42}(\aaddr))} \in \lrv(W_{R,\aaddr}')
+    \]
+    for $n' < n-1$.
     If we can show
     \[
-      \npair[n-1]{(\Phi_S.\ms_\stk(\aaddr),\ms_\var{T,free\_stack}(\aaddr))} \in \lrv(W_{R,\aaddr})
-    \]
-    then we are done by monotonicity of $\lrv$.
+      \npair[n']{(\Phi_S.\ms_\stk(\aaddr),\ms_\var{T,free\_stack}(\aaddr))} \in \lrv(W_{R,\aaddr})
+    \]
+    for $n' < n-1$, then we are done by monotonicity of $\lrv$.
 
@@
     \[
-      \npair{R_\ms(r)} \in W_{R,\rstk}(r).H \; \xi^{-1}(R_W(r))
-    \]
+      \npair[n']{R_\ms(r)} \in W_{R,\rstk}(r).H \; \xi^{-1}(R_W(r))
+    \]
+    for $n' < n$
     which entails
@@
 
-    Now this is exactly what we wanted to show because $R_W(r)$ is what we picked as $W_{R,\aaddr}$.
-  \item $\memSatFStack{\Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]},\ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}}{W_{M,\var{free\_stack}}'}$:
+    Now, using Lemma~\ref{lem:downwards-closed}, this is exactly what we wanted to show because $R_W(r)$ is what we picked as $W_{R,\aaddr}$.
+  \item $\memSatFStack[n-1]{\Phi_S.\ms_\stk - \Phi_S.\ms_\stk|_{[\aaddr_\stk,\eaddr_\stk]},\ms_\var{T,free\_stack} \setminus \ms_\var{T,free\_stack}|_{[\aaddr_\stk,\eaddr_\stk]}}{W_{M,\var{free\_stack}}'}$:
 
@@
 
-  \item $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap[W_M'])(W_{M,\var{heap}}')$:
-
-    Follows from Lemma~\ref{lem:heap-sat-unchanged-future}, the fact the the heap part of the world remains unchanged and $W_{M,\var{heap}}' \future W_{M,\var{heap}}$.
+  \item $\npair[n-1]{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap[W_M'])(W_{M,\var{heap}}')$:
+
+    Follows from Lemma~\ref{lem:heap-sat-unchanged-future}, Lemma~\ref{lem:downwards-closed}, the fact that the heap part of the world remains unchanged and $W_{M,\var{heap}}' \future W_{M,\var{heap}}$.
   \end{itemize}

### 2018-07-03 fully-abstract-comp/technical-report.tex
Fix non-expansiveness of stack, free stack and heap satisfaction, adapt step accounting throughout the paper to follow suit.

@@
       \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
-      \qquad \forall r \in \dom(\activeReg{W_\var{stack}})\ldotp \\
-      \qquad \quad \npair{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n\ldotp \\
+      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
@@
       \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
-      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}) \ldotp \\
-      \qquad \quad \npair{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}),n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
     \end{array}
@@
       \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
-      \qquad \forall r \in \dom(\activeReg{\pwheap}) \ldotp \\
-      \qquad \quad \npair{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
       \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
@@
   \lrheap(\pwheap)(W_{\var{heap}})$, we get an
-  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $\npair{R_\ms(r_{\var{code}})} \in  \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$.
+  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $\npair[n']{R_\ms(r_{\var{code}})} \in  \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$ for all $n' < n$.
 
@@
       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-      &\qquad\quad\text{then } \npair[n']{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \text{ for all $n' < n$} \tand\\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
@@
   \begin{equation*}
-    \npair[n']{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
   \end{equation*}
-  for all $n' < n$ and
+  and
   \begin{equation*}
@@
   \begin{align*}
-    \npair[n']{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1]', \mscode[1]\uplus\msdata[1]')} &\in \lrheap(\pwheap[W_1])(W')\\
-    \npair[n']{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2]', \mscode[2]\uplus\msdata[2]')} &\in \lrheap(\pwheap[W_2])(W')
+    \npair{(\sigrets[1]\uplus\sigcloss[1],\mscode[1]\uplus \msdata[1]', \mscode[1]\uplus\msdata[1]')} &\in \lrheap(\pwheap[W_1])(W')\\
+    \npair{(\sigrets[2]\uplus\sigcloss[2],\mscode[2]\uplus \msdata[2]', \mscode[2]\uplus\msdata[2]')} &\in \lrheap(\pwheap[W_2])(W')
   \end{align*}
-  for all $n' < n$.
   It follows by Lemma~\ref{lem:combined-memory-disjoint-world} that 
   \begin{equation*}
-    \npair[n']{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
+    \npair{(\sigrets[3]\uplus\sigcloss[3],\mscode[3]\uplus \msdata[3]', \mscode[3]\uplus\msdata[3]')} \in \lrheap(\pwheap[(W_1\uplus W_2)])(W')
   \end{equation*}
-  for all $n' < n$.
 
@@
     \begin{equation*}
-      \npair[n']{(\sigrets\uplus\sigcloss,\mspad\uplus\mscode\uplus \msdata', \mspad\uplus\mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
+      \npair{(\sigrets\uplus\sigcloss,\mspad\uplus\mscode\uplus \msdata', \mspad\uplus\mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
     \end{equation*}
-    for all $n' < n$.
 
@@
   \begin{equation*}
-    \npair[n']{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
+    \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W')
   \end{equation*}
-  for all $n' < n$ and
+  and
   \begin{equation*}
@@
     \begin{equation*}
+      \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap[W_M])(W_M)
+    \end{equation*}
+    and by Lemma~\ref{lem:downwards-closed} also
+    \begin{equation*}
       \npair[n']{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata, \mscode\uplus\msdata)} \in \lrheap(\pwheap[W_M])(W_M)
@@
     \begin{equation*}
-      \npair[n']{(\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})} \in \stdreg{\{a\},\gc}{\spatialo}.H~ \xi^{-1}(\purePart{W_M})
+      \npair[n'']{(\ms_\stk|_{\{a\}},\ms_\stk|_{\{a\}})} \in \stdreg{\{a\},\gc}{\spatialo}.H~ \xi^{-1}(\purePart{W_M})
     \end{equation*}
+    for all $n'' < n'$.
     By definition, it suffices to show that for all $n'' < n'$, we have that:

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Reformulation and typo.

--- 2018-07-02 fully-abstract-comp/popl2019/paper.tex
@@
 One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
-The callee would have to give up this token on return to prevent it from spending the same token multiple times As it turns out, the stack capability for the unused part of the stack can be seen as such a token: When the callee returns, we expect that it returns the stack capability we gave it in the call.
+The callee would have to give up this token on return to make sure that the token is only spent once.
+As it turns out, the stack capability for the unused part of the stack can be seen as such a token: When the callee returns, we expect that it returns the stack capability we gave it in the call.
 If this stack capability is not present, then we have no whether our stack frame is on top of the stack, so we should refuse to continue execution.

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Added a new figure and restructured document

@@
 \section{Linear Stack and Return Capabilities}
-\todo[inline]{The stack grows downwards, maybe we should make the figures reflect this.}
+\todo[inline]{The stack grows downwards, maybe we should make the figures reflect this. Also make sealed capabilities more clear by greying out.}
 \begin{figure}
@@
 The callee would have to give up this token on return to make sure that the token is only spent once.
-As it turns out, the stack capability for the unused part of the stack can be seen as such a token: When the callee returns, we expect that it returns the stack capability we gave it in the call.
-If this stack capability is not present, then we have no whether our stack frame is on top of the stack, so we should refuse to continue execution.
-If we receive a stack pointer, then we try to splice it with the stack pointer for our local stack frame (which by now has been unsealed as we resume execution).
-The splice instruction only splices adjacent capabilities, so in the case where the stack frame that is returned to is on top of the stack, the returned stack capability and the unsealed stack capability are adjacent, so the splice is successful.
-In the case where the stack frame is not on top of the call stack, as illustrated in Figure~\ref{fig:ret-ptr-abuse}, part of the stack capability will be bound in the sealed return pairs for the higher stack frames making the splice fail as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}.
-
-% Where do we get the stack pointer from, how can we know it is linear
-% The other attack.
-
-% Non-empty trusted stack frames.
-
-% known stack base
-% + more
-% non-empty stack frames
-
+As it turns out, the stack capability for the unused part of the stack can be seen as such a token, so when the callee returns, we expect them to return the stack capability we gave them in the call.
+When we receive a stack capability on return, we will try to splice it with the now unsealed stack capability for our private stack frame.
+If this is successful, then we know that the two capabilities are adjacent.
+If it is unsuccessful or if no stack capability is returned, then we cannot determine whether our stack frame is the topmost.
+The issue illustrated in Figure~\ref{fig:ret-ptr-abuse} is prevented by splicing as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}.
 \begin{figure}
@@
 
+% Non-empty trusted stack frames.
+
+% Where do we get the stack pointer from, how can we know it is linear
+% The other attack.
+
+% known stack base
+% + more
+% non-empty stack frames
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.2\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \actadv{(0,4)}{(4.5,7)}{}
+
+      \lincapbrace{(4.6,2)}{(4.6,4)}{}
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+      \lincapbracebot{(4.6,4)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-a}
+  \end{subfigure}
+  \begin{subfigure}{0.24\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \inactadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
+      \node[opacity=0.5,rotate=-45] at (2.25,10) {\footnotesize Sent to trusted};
+      \lincapbrace{(4.6,2)}{(4.6,4)}{}
+      \lincapbrace{(4.6,4)}{(4.6,7)}{}
+
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+      \lincapbracebot{(4.6,7)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-b}
+  \end{subfigure}
+  \begin{subfigure}{0.24\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \actadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
+      \node[opacity=0.5,rotate=-45] at (2.25,11) {\footnotesize Returned to adv.};
+      \lincapbrace{(4.6,2)}{(4.6,4)}{}
+      \lincapbrace{(4.6,4)}{(4.6,7)}{}
+      \lincapbrace{(4.6,7)}{(4.6,9)}{}
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-c}
+  \end{subfigure}
+  \begin{subfigure}{0.24\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \actsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
+      \begin{scope}
+        \clip (-.1,-.1) rectangle (4.6,13.1);
+        \draw[fill=gray!50] (0,9) rectangle (4.5,13.5);
+      \end{scope}
+
+      \lincapbrace{(4.6,2)}{(4.6,7)}{}
+      \lincapbrace{(4.6,7)}{(4.6,9)}{}
+
+      \fill[red,draw=red,opacity=.4] (-.1,6.9) rectangle (4.6,9.1);
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-d}
+  \end{subfigure}
+
+  \label{fig:stk-base-abuse}
+\end{figure}
+
 \begin{itemize}

### 2018-07-03 fully-abstract-comp/popl2019/preamble-illu.tex
Added a new figure and restructured document

@@
 
+\newcommand{\lincapbracebot}[4][sp1]{
+  \draw [densely dotted,decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt,aspect=0.25},yshift=0pt]
+  #2 -- #3 node (#1) [black,near start,xshift=1cm,align=center] {\footnotesize #4};
+}
+
 \newcommand{\capbracebotred}[4][sp1]{
@@
 
+\newcommand{\stdstackstartclean}[1][11]{
+  \scope
+    \clip (-.1,-.1) rectangle (4.6,#1 + .1);
+    \fill[fill=white] (0,0) rectangle (4.5,#1);
+    \draw (0,0) -- (0,#1);
+    \draw (4.5,0) -- (4.5,#1);
+  \endscope
+  % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+  \draw (-1.5,0) node {};
+  \draw (-1.5,#1 + 1) node {};
+}
+
 \newcommand{\stdstackstart}[1][11]{
@@
     \draw (4.5,0) -- (4.5,#1);
-    \draw[fill=gray!50] (0,-.5) rectangle (4.5,2)
-    node[pos=.5,color=black,align=center,yshift=0.1cm] {\footnotesize Lower
-      stack\\\footnotesize frames...};
+    \draw[fill=gray!50] (0,-.5) rectangle (4.5,2)node[pos=.5,color=black,align=center,yshift=0.1cm] {\footnotesize Lower stack\\\footnotesize frames...};
   \endscope

### 2018-07-03 fully-abstract-comp/ftlr-proof-cases.tex
Argue categorisation of operational semantics steps.

@@
     \end{itemize}
-  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\normal),\baddr,\eaddr,\_)$
+  \item $\Phi_S(\pcreg) = \Phi_T(\pcreg) = ((\rx,\normal),\baddr,\eaddr,a)$
   \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$
@@
     \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r}{z}} \neq \failed$
+    \item $r \neq \pcreg$
     \item $z \in \ints$
@@
     \item $\Phi_T \step \Phi_T' \neq \failed$
-      % domi: this condition doesn't make syntactic sense and is also not used.
-    % \item $\Phi_T' = \updPc{\Phi_T\updReg{\rtmp{1}}{z}}$
-    \item $z \in \ints$
     \item $\Phi_S(r_1) = \sealed{\sigma,c_1}$
@@
         \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $k = 1$
-      \item \label{case:ftlr:move-cca-etc:move}(move) $w_1 = \Phi_S(r_2)$, $w_1' = \Phi_T(r_2)$, and $\isLinear{w_1} \Rightarrow w_2 = w_2' = 0$, and $\neg \isLinear{w_1} \Rightarrow (w_2 = \Phi_S(r_2) \wedge w_2' = \Phi_T(r_2)$) and $r_1 \neq r_2$ and $k = 2$.
-      \item \label{case:ftlr:move-cca-etc:move-eq}(move) $w_1 = \Phi_S(r_1)$, $w_1' = \Phi_T(r_1)$ and $k = 1$.
-      \item \label{case:ftlr:move-cca-etc:cseal}(cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' = \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \le \sigma \le \sigma_\eaddr$, and $k=1$, and either $\Phi_S$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ or $\ta \# [\baddr,\eaddr]$.
+      \item \label{case:ftlr:move-cca-etc:move}(move) $w_1 = \Phi_S(r_2)$, $w_1' = \Phi_T(r_2)$, and $w_2 = \linCons{w_1}$, $w_2' = \linCons{w_1'}$ and $r_1 \neq r_2$ and $k = 2$.
+      \item \label{case:ftlr:move-cca-etc:cseal}(cseal) $w_1 = \sealed{\sigma,\Phi_S(r_1)}$, $w_1' = \sealed{\sigma,\Phi_T(r_1)}$, $\Phi_S(r_2) = \Phi_T(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, $\sigma_\baddr \le \sigma \le \sigma_\eaddr$, and $k=1$, and $\Phi_S$ points to $\src{\tcseal{r_1}{r_2}}$.
       \item \label{case:ftlr:move-cca-etc:split-normal} (split) $\Phi_T(r_3) = \Phi_S(r_3) =
@@
       \item \label{case:ftlr:move-cca-etc:split-stack} (split) $\Phi_S(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$,
-        $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq s$,
-        $s < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' =
-        ((\perm,\linear),\baddr,s,\aaddr)$, $w_2 =
-        \stkptr{\perm,s+1,\eaddr,\aaddr}$,
-
-        $w_2' = ((\perm,\linear),s+1,\eaddr,\aaddr)$, $k=2$.
+        $\Phi_T(r_3) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$, $\baddr \leq n$,
+        $n < \eaddr$, $w_1 = \stkptr{\perm,\baddr,n,\aaddr}$, $w_1' =
+        ((\perm,\linear),\baddr,n,\aaddr)$, $w_2 =
+        \stkptr{\perm,n+1,\eaddr,\aaddr}$,
+
+        $w_2' = ((\perm,\linear),n+1,\eaddr,\aaddr)$, $k=2$.
       \item \label{case:ftlr:move-cca-etc:splice-stack}(splice) $\Phi_T(r_2) = ((\perm,\linear),\baddr_2,\eaddr_2,\_)$,
@@
           \sigma_{\eaddr,3}, \sigma}$
-      \item \label{case:ftlr:move-cca-etc:jnz-z} (jnz zero case) $k = 0$
+      \item \label{case:ftlr:move-cca-etc:jnz-z} (jnz zero case, noop move) $k = 0$
       \end{enumproof}
@@
         $w = \Phi_S(r_2)$, and $w' = \Phi_T(r_2)$, and $w_2 = \linCons{\Phi_S(r_2)}$, $w_2' = \linCons{\Phi_T(r_2)}$.
-      \item (load) $ w_2' = \Phi_T(r_2) =
+      \item (load) $w_2' = \Phi_T(r_2) =
         ((\perm,\linear),\baddr,\eaddr,\aaddr)$, and $w_2 = \Phi_S(r_2) =
@@
         \dom(\Phi.\ms_\stk)$, and $\aaddr \in \dom(\Phi.\ms_\stk)$, and\\
-        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm,w_1}$ and $\linConsPerm{\perm,w_1'}$.
+        $w_1 = \Phi_S.\ms_\stk(\aaddr)$, and $w_1' = \Phi_T.\ms_\stk(\aaddr)$, and $w = \linCons{w_1}$, $w' = \linCons{w_1'}$, $\linConsPerm{\perm}{w_1}$ and $\linConsPerm{\perm}{w_1'}$.
       \end{enumproof}
     \end{itemize}
-  \item \label{case:ftlr:store-load-jump} All of the following hold: (includes control-flow manipulation cases:
+  \item \label{case:ftlr:jump} All of the following hold: (includes control-flow manipulation cases:
     jmp, jnz, xjmp, that can be handled mostly uniformly)
@@
   \end{enumproof}
-  \lau{TODO: Argue that the above is indeed the case.}
+  The above follows from a careful analysis of the cases of the operational semantics, using the following facts:
+  \begin{itemize}
+  \item $[b,e] \subseteq \ta$ or $[b,e] \mathrel{\#} \ta$ (by assumption of this lemma)
+  \item $\Phi_S.\reg(\pcreg) = \Phi_T.\reg(\pcreg)$ (by assumption of this lemma)
+  \item $\Phi_S.\mem(a) = \Phi_T.\mem(a)$ if $a \in [b,e]$: follows from
+    $\npair{[\baddr,\eaddr]} \in \xReadCond{W_\pcreg}$, the fact that $W_\pcreg
+    \oplus W_R \oplus W_M$ is defined and
+    $\memSat{\Phi_S.\mem,\Phi_S.\stk,\Phi_S.\ms_\stk,\Phi_T.\mem}{W_M}$
+  \item $\Phi_S.\mem(a\cdots a+\calllen-1) = \Phi_T.\mem(a \cdots a+\calllen -1)$ if $[a\cdots a+\calllen-1] \subseteq [b,e]$ (follows similarly).
+  \item $\Phi_S.\reg(r) = ((\perm_1,\lin_1),\baddr_1,\eaddr_1,\aaddr_1)$ implies that $\Phi_S.\reg(r) = \Phi_T.\reg(r)$ and $a_1 \in \dom(\Phi_S.\mem)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) = z$ implies that $\Phi_S.\reg(r) = \Phi_T.\reg(r)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) = \seal{\sigma_b,\sigma_e,\sigma}$ implies that $\Phi_S.\reg(r) = \Phi_T.\reg(r)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) = \sealed{\sigma,\vsc}$ implies that $\Phi_T.\reg(r) = \sealed{\sigma,\vsc'}$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) \neq \retptrd{\_,\_}$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) \neq \retptrc{\_,\_}$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) = \stkptr{\perm_1,\baddr_1,\eaddr_1,\aaddr_1}$ implies that $\Phi_T.\reg(r) = ((\perm_1,\linear),b_1,e_1,a_1)$ and $a \in \dom(\Phi_S.\ms_\stk)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item Similar facts about the current address, base and end pointer, permissions and linearity of all register capabilities being equal (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \end{itemize}
 
@@
     which follows from Lemma~\ref{lem:monotonicity} and \ref{lem:world-fut-purePart} and what we have proven about the pc.
-  \item Case \ref{case:ftlr:move-cca-etc:move-eq}:
-
-    Follows from Assumption and Lemma~\ref{lem:downwards-closed}.
-
   \item Case \ref{case:ftlr:move-cca-etc:cseal}:
@@
 
-  Case \ref{case:ftlr:store-load-jump}
+  Case \ref{case:ftlr:jump}
 

### 2018-07-03 fully-abstract-comp/technical-report.tex
Argue categorisation of operational semantics steps.

@@
 \end{align*}
-\dominique{9-3-2018: $\linCons$ should only be allowed to modify the memory if the capability has write permission!}
 
@@
 If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
-\lau{28-09-2017: For now, the jump to a return pointer does the stack end point check. We could possibly move this to the proposed semantic condition as well.}
-% \begin{align*}
-%   \sem{\txjmp{r_1}{r_2}}(\Phi) = &
-%                                    \begin{cases}
-%                                      \arraycolsep=0pt
-%                                      \begin{array}[t]{rl}
-%                                        \Phi & \updReg{r_1}{w_1}\\
-%                                             & \updReg{r_2}{w_2}\\
-%                                             & \updReg{\pcreg}{c_\code}\\
-%                                             & \updReg{\rdata}{c_\data}
-%                                      \end{array} &
-%                                      \begin{array}[t]{l}
-%                                        \Phi(r_1) = \sealed{\sigma_1,c_\code} \tand \\
-%                                        \Phi(r_2) = \sealed{\sigma_2,c_\data} \tand \\
-%                                        \sigma_1 = \sigma_2 \tand \\
-%                                        \nonExec{c_\data} \tand \\
-%                                        w_1 = \linCons{\Phi(r_1)} \tand \\
-%                                        w_2 = \linCons{\Phi(r_2)}
-%                                      \end{array}
-%                                      \\&\\
-%                                      \sourcecolor\left.
-%                                        \arraycolsep=0pt
-%                                        \begin{array}[t]{rl}
-%                                          \Phi' & \updReg{r_2}{0} \\
-%                                                & \updReg{\pcreg}{\opc}\\
-%                                                & \updReg{\rdata}{0}\\
-%                                                & \updReg{\rstk}{c_\stk}\\
-%                                                & \updReg{\rtmp{1}}{0}\\
-%                                                & \updReg{\rtmp{2}}{0}
-%                                        \end{array}\right.&
-%                                      \sourcecolor\left.
-%                                        \begin{array}[t]{l}
-%                                          \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
-%                                          \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
-%                                          \sigma_1 = \sigma_2 \tand\\
-%                                          \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
-%                                          \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
-%                                          \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
-%                                          \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
-%                                          \eaddr_{\stk} + 1 = \aaddr_\stk \tand\\
-%                                          c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
-%                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
-%                                        \end{array}\right.
-%                                      \\
-%                                      \failed & \totherwise
-%                                    \end{cases}
-% \end{align*}
-
-% \dominique{22-1-2018: the below is an alternative proposal to the above definition.
-It is part of a proposed refactoring to simplify the merge the cases for return pointers and closures in the value relation.}
 \begin{align*}
@@
 
-\lau{19-09-2017: Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
+Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
 This means that if a caller tries to use a return pointer with a stack that is not immediately adjacent to the private stack, then it fails.
 (2) Splice prohibit splicing with an empty capability!
-This means that a callee cannot return an empty stack (this also means that it is impossible to make a call when all of the stack is used - this may indeed be undesirable, but without this restriction we need to handle other things).}
-\lau{19-09-2017: Because $\tsplice{}{}{}$ does not allow empty stacks, it is not ``left inverse'' to $\tsplit{}{}{}{}$ (because of the empty case).
-Intuitively, it is weird that a $\tsplit{}{}{}{}$ followed by a $\tsplice{}{}{}$ does not yield the same capability.}
+This means that a callee cannot return an empty stack (this also means that it is impossible to make a call when all of the stack is used - this may indeed be undesirable, but without this restriction we need to handle other things).
+Note: because $\tsplice{}{}{}$ does not allow empty stacks, it is not ``left inverse'' to $\tsplit{}{}{}{}$ (because of the empty case).
+Intuitively, it is weird that a $\tsplit{}{}{}{}$ followed by a $\tsplice{}{}{}$ does not yield the same capability.
 \begin{align*}
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
     \; \middle| \;

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex

@@
 \end{itemize}
+
 \section{A Capability Machine with Sealing and Linear Capabilities}
+\label{sec:capab-mach-with}
 \begin{itemize}
@@
 \end{itemize}
+
 \section{Formulating Security with a Fully Abstract Overlay Semantics}
+\label{sec:form-secur-with}
 \begin{itemize}
@@
 \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
+\item Performance?  Implementability of linear caps?
 \end{itemize}
@@
 \section{Related Work}
+
+Capability machines originate with Dennis and Van Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to Levy~\cite{Levy1984capability} and Watson et al.~\cite{Watson2015Cheri} for an overview of previous work.
+The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is modeled CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
+The latter is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
+In the interest of simplicity, we have omitted features of CHERI that were not needed for modelling StackTokens (e.g.\ local capabilities, virtual memory).
+
+Plenty of other papers enforce well-bracketed control flow at a low level, but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+This includes particularly the line of work on \emph{control-flow integrity}~\cite{abadi_control-flow_2005}.
+This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
+Contrary to StackTokens, CFI can be implemented on commodity hardware rather than capability machines.
+However, its attacker model is different and its security goals are weaker.
+They assume an attacker that is not able to execute code, but can overwrite arbitrary data at any time during execution (to model buffer overflows).
+In terms of security goals, the technique does not enforce local stack encapsulation.
+Also, it only enforces a weak form of control flow correctness, saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
+Such a property ignores temporal properties and seems hard to use for reasoning.
+
+More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}.
+Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
+TODO: adapt the following sentence...
+Also, both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, while we define a logical relation that can be used to reason about specific programs that rely on well-bracketed control flow.
+
 \begin{itemize}
 \item CFI and friends
+\item CheriBSD etc.
 \item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
 \item undefined behavior and full abstraction \citep{juglaret_beyond_2016} \dominique{TODO}
-\item CheriBSD etc.
 \end{itemize}

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Some work on related work.

@@
 \section{Introduction}
+\label{sec:introduction}
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
 More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
-These processors use tagged memory to enforce a strict separation between integers and \emph{capabilities}: pointers that carry authority.
+These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
 Capabilities come in different flavours.
 Memory capabilities allow reading from and writing to a block of memory.
-Additionally, capability machines offer some form of \emph{object capabilities} that represent low-level encapsulated closures, i.e.
+Additionally, capability machines offer some form of {\itshape object capabilities} that represent low-level encapsulated closures, i.e.
 a piece of code coupled with private state that it gains access to upon invocation.
@@
 While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
-For example, CHERI offers a form of \emph{local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
+For example, CHERI offers a form of {\itshape local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
 \citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and a number of security checks and measures, the two properties can be guaranteed.
@@
 In this work, we propose and study StackTokens: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on \emph{linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
 These capabilities are enforced by the hardware to be non-duplicable.
@@
 
-A second contribution of this work is the way in which we formulate these two properties, using a technique we call \emph{fully abstract overlay semantics}.
+A second contribution of this work is the way in which we formulate these two properties, using a technique we call {\itshape fully abstract overlay semantics}.
 Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
@@
 \item Performance?  Implementability of linear caps?
+\item Sharing stack references accross component boundaries is supported
 \end{itemize}
@@
 \section{Related Work}
+
+In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.
+We do not repeat the work we discussed in Section~\ref{sec:introduction}.
 
@@
 Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
-TODO: adapt the following sentence...
-Also, both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, while we define a logical relation that can be used to reason about specific programs that rely on well-bracketed control flow.
+Both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
 
-\begin{itemize}
-\item CFI and friends
-\item CheriBSD etc.
-\item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
-\item undefined behavior and full abstraction \citep{juglaret_beyond_2016} \dominique{TODO}
-\end{itemize}
+CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
+The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation at play.
+They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared accross component boundaries.
+
+The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
+In their study of compartmentalization primitives, \Cite{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that is able to deal with source languages with undefined behavior (together with other benefits).
+Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
+In follow-up work, they extend this approach to apply in scenario's where components only exhibit undefined behavior after a number of well-defined steps~\citep{DBLP:journals/corr/abs-1802-00588}.
+If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
+
 \bibliography{references}

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Fix citations.

@@
 
-Capability machines originate with Dennis and Van Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to Levy~\cite{Levy1984capability} and Watson et al.~\cite{Watson2015Cheri} for an overview of previous work.
-The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is modeled CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
+Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
+The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is modeled CHERI~\cite{watson_cheri:_2015,woodruff_cheri_2014}.
 The latter is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
@@
 
-More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}.
+More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016,juglaret_beyond_2016}.
 Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
@@
 The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
-In their study of compartmentalization primitives, \Cite{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that is able to deal with source languages with undefined behavior (together with other benefits).
+In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that is able to deal with source languages with undefined behavior (together with other benefits).
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Tune the related work discussion.

@@
 Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
-The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is modeled CHERI~\cite{watson_cheri:_2015,woodruff_cheri_2014}.
-The latter is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
-In the interest of simplicity, we have omitted features of CHERI that were not needed for modelling StackTokens (e.g.\ local capabilities, virtual memory).
+The capability machine formalized in Section~\ref{sec:capab-mach-with} is modeled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
+This is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
+For simplicity, we have omitted features of CHERI that were not needed for modelling StackTokens (e.g.\ local capabilities, virtual memory).
 
 Plenty of other papers enforce well-bracketed control flow at a low level, but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
-This includes particularly the line of work on \emph{control-flow integrity}~\cite{abadi_control-flow_2005}.
+This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
 This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
@@
 The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
-In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that is able to deal with source languages with undefined behavior (together with other benefits).
+In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that applies to unsafe source languages.
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, they extend this approach to apply in scenario's where components only exhibit undefined behavior after a number of well-defined steps~\citep{DBLP:journals/corr/abs-1802-00588}.
+In follow-up work, \citet{DBLP:journals/corr/abs-1802-00588} extend this approach to scenario's where components only start to exhibit undefined behavior after a number of well-defined steps.
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
+Our results do not apply to dynamic compromise scenario's because they are intended to be used in the verification of a secure compiler, where these scenario's are not relevant.
 

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Add a note about the limitations of CFI.
typos

@@
 Such a property ignores temporal properties and seems hard to use for reasoning.
+There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending}.
 
@@
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, \citet{DBLP:journals/corr/abs-1802-00588} extend this approach to scenario's where components only start to exhibit undefined behavior after a number of well-defined steps.
+In follow-up work, \citet{DBLP:journals/corr/abs-1802-00588} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps.
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenario's because they are intended to be used in the verification of a secure compiler, where these scenario's are not relevant.
+Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler, where these scenarios are not relevant.
 

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Text justifying need to be able to check entire stack.

@@
   \end{subfigure}
-
+  \caption{}
 \end{figure}
@@
 % Non-empty trusted stack frames.
+In this token-based scheme, it is important that the caller keeps a bit of the token which corresponds to having a non-empty stack frame.
+In a traditional C-like calling convention, this is natural as the caller needs to push information such as the old stack pointer and the old program counter to the stack.
+On a capability machine with CHERI's sealed capabilities, this is not necessary as the old program counter capability and the capability for the caller's stack frame is sealed in the return pair, and they will become unsealed when jumped to.
+If the caller has an empty stack frame, then they have entirely given up all of their return token which practically means that they are not present on the call stack and thus have no way verify whether it is their turn to be returned to.
+
+% known stack base
+% + more
+We check that the returned stack capability is adjacent to the stack capability for our local stack frame, but this does not guarantee that it is the same token as we gave away in the call.
+Specifically, we have no idea whether the callee has kept part of the token for themselves and only given us back the part that is adjacent to our token.
+If we do not make sure that the entire token is returned, then it is possible for an adversary to break WBCF.
+This is illustrated in Figure~\ref{fig:stk-base-abuse}: Some trusted code has created a stack frame with their local data and called a piece of untrusted code giving them access to the remainder of the stack.
+This situation is illustrated in Figure~\ref{fig:stack-base-abuse-a}.
+Now the adversarial code keeps the stack capability for the part of the stack adjacent to the trusted code's stack frame.
+The adversary does not seal this part in a return pair, but simply stores it on the heap.
+At this point the adversary calls the trusted code with the remainder of the stack capability as illustrated in Figure~\ref{fig:stack-base-abuse-b}.
+At some point the trusted code calls the adversary again leaving a second stack frame on the stack an giving the adversary the rest.
+At this point, the adversary has access to the stack capability passed in the second call performed by the trusted code as well as the stack capability stored away after the first call (Figure~\ref{fig:stack-base-abuse-c}).
+Now the adversary has access to a stack capability for part of the stack adjacent to the first stack frame of the trusted caller.
+The adversary can use this stack capability to return from the first call without ever returning from the second call (Figure~\ref{fig:stack-base-abuse-d}).
 
@@
 
-% known stack base
-% + more
-% non-empty stack frames
+% Summary of CC
 
@@
   \centering
-  \begin{subfigure}{0.2\linewidth}
+  \begin{subfigure}{0.23\linewidth}
     \centering
@@
   \end{subfigure}
-
+  \caption{}
   \label{fig:stk-base-abuse}

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Added arrow to alert reader to the address space layout in figures.
Finished text for example for check base.

@@
       % recurrent parts
+      \draw (-0.5,13) node {\footnotesize 0};
+      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
       \stdstackstart[13]
@@
 Now the adversary has access to a stack capability for part of the stack adjacent to the first stack frame of the trusted caller.
-The adversary can use this stack capability to return from the first call without ever returning from the second call (Figure~\ref{fig:stack-base-abuse-d}).
+The adversary can use this stack capability to return from the first call without ever returning from the second call thus breaking WBCF (Figure~\ref{fig:stack-base-abuse-d}).
+In this scenario, the issue arises because we do not check whether the entire token was returned to us.
+We could do this by checking whether the base address\footnote{The stack grows downwards in the address space} of the token returned to us correspond to the base address of the token we provided in the call.
+In the scenario we just sketched, we would be alerted to an attempt to break WBCF as the base address of the token returned to us in Figure~\ref{fig:stack-base-abuse-d} does not correspond to the base address of the token we passed to the adversary in Figure~\ref{fig:stack-base-abuse-a}.
 

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
First draft of section 2 (linear stack and return capabilities)

@@
 \section{Linear Stack and Return Capabilities}
-\todo[inline]{The stack grows downwards, maybe we should make the figures reflect this. Also make sealed capabilities more clear by greying out.}
+\todo[inline]{Make sealed capabilities more clear by greying out.}
 \begin{figure}
@@
 
-% informally explain how we prevent this using linear capabilities
-\lau[inline]{Write an introductory paragraph to the linear capability bit}
+% Informally explain how we prevent this using linear capabilities
+As the examples illustrate, the standard memory and object-like capabilities do not provide sufficient guarantees to enforce LSE and WBCF.
+The two properties can be enforced with a calling convention that utilizes the guarantees of linear capabilities.
+In the following, we will describe such a calling convention through motivating examples, and in later sections, we will prove that it actually enforces LSE and WBCF.
+We call the calling convention StackTokens for reasons that will become clear later.
 
@@
 % The other attack.
+At this point, we have the means to check whether a token actually corresponds to the one that we passed away.
+So far, we have claimed that this can be used to make certain that it is our turn on some call stack, but in order to have WBCF all the well-behaved code must use the same call stack.
+In the case where we are invoked by some adversarial code, we have no way to be sure whether the stack capability passed to us stems from the same stack used to call other well-behaved code.
+For example, an adversary could call us with a callback and some linear capability as the stack capability.
+Assuming we choose to trust the stack, we use it for the necessary token and invoke the callback.
+Now, the adversary could take a completely different linear capability and call some well-behaved code (it could be us) with a callback.
+The same thing happens again: The callback is called, so a return token is generated and passed to the adversary.
+At this point, the adversary has a return token from passed in each call of the callback.
+Both of the return tokens are intact which means that the adversary can return from either callback.
+From a high-level perspective, the problem is that both well-behaved programs believe that they are on top of the call stack.
+They are both correct as the adversarial code has created two different call stacks.
+In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
+We do this by deciding on a global base address for the stack which is used for the entirety of the execution.
+In order to make sure that we are actually using the correct stack, we check that the base address of the stack capability we receive corresponds to the globally decided stack base address.
+This also means that on every return, it will be this address that we check for when we make sure that the entire token was returned to us.
+
+Having a globally decided stack address also solves another challenge.
+Namely, the stack capability has to be linear in order for us to use it as a token, so in principle we would have to check that this is the case whenever we receive a stack capability.
+This is, however, checked when we check that the stack capability governs the global stack base address.
+The stack capability needs to be linear for us to use it as a token, and if the stack base address is part of the stack, then it must be governed by a linear capability when execution begins. 
+Due to the non-duplicability of linear capability and the fact that they cannot become non-linear during execution, we know during execution that if we have a capability that governs the stack base address, then it must be linear.
+
+% Return pointers/return seals
+The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter and the data part is the stack capability for the local stack frame of the caller.
+Both of the capabilities in the pair are of course sealed with the same seal.
+% One seal per return point
+All call points need to be associated with a unique seal that is only used to seal the capabilities in the return pair for that call point.
+We refer to the seals used to create return pairs as return seals.
+The return seal is what associates the stack frame on the call stack with a call in a program, so if we allowed return seals to be reused, it would be possible to return to a different call than the one that caused the stack frame to be pushed on the call stack which would break WBCF.
+For similar reasons, we cannot allow return seals to be used to seal a closure.
+% Don't be stupid
+Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
+This goes for direct leaks, such as leaving a seal in a register or writing it to adversarial memory, as well as indirect leaks, where a capability for reading a return seal from memory is leaked (either directly or indirectly).
+
+% Note about them vs us, we do nothing that they couldn't do
+In the description of StackToken, we have described it in terms of ``them vs us''.
+This may create the impression of an asymmetric calling convention that allows us to protect us against some adversary, but what if the adversary does not trust us and want to guarantee LSE and WBCF?
+The answer is simple: We have made no assumptions that the adversary cannot make, and we put no restrictions on the adversary that we do not follow, so they can employ StackTokens to get the exact same guarantees as us!
 
 % Summary of CC
+In the previous section, we have described our proposed calling convention, StackTokens.
+In this paragraph we summarize it:
+\begin{description}
+\item[Check the base address of the stack capability before and after calls] 
+\item[Make sure that local stack frames are non-empty]
+\item[Create token and data return capability on call] split the stack capability in two to get a stack capability for you local stack frame and a stack capability for the unused part of the stack.
+The stack capability for your local stack frame is sealed and used for the data part of the return pair.
+\item[Create code return capability on call] Seal the old program pointer.
+\item[Reasonable use of seals] return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
+\end{description}
 

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Pass over section 2.

--- 2018-07-03 fully-abstract-comp/popl2019/paper.tex
@@
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
-In both examples some trusted and adversarial code has been taking turns calling each other, so the stack now alternates between stack frame of the trusted and adversarial code.
-The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how the trusted code tries to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
-However, restricting the stack capability does not help when the trusted code in the first call gave access to the part of the stack where the stack frame of the second call now resides.
-Generally speaking, the trusted code has no reason to ever trust a stack capability it receives from adversarial code as the adversarial code may have duplicated the stack capability and stored it for later use.
-In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), the trusted code has given the adversarial code two sealed pairs of return capabilities, one in each of the calls to the adversarial code.
-The adversarial code may have stored the pair of sealed return pointers from the first call, so that it can access it in the second call.
-The figure illustrates how the adversarial code uses the return pair from the first call breaking WBCF.
+In both examples our component and some adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
+The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
+However, restricting the stack capability does not help when we in the first call give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
+Generally speaking, we have no reason to ever trust a stack capability received from a component we do not trust as that stack capability may have been duplicated and stored for later use.
+In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
+The adversary may have stored the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed to use it.
+The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call an thus breaks WBCF.
 
-% informally explain how we prevent this using linear capabilities
-\lau[inline]{Write an introductory paragraph to the linear capability bit}
-
+\todo[inline]{If target cap. machine is introduced before this section, then revisit/delete this section.}
+% Informally explain how we prevent this using linear capabilities
+As the examples illustrate, the standard memory and object-like capabilities (sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
+The two properties can be enforced with a calling convention that utilizes the guarantees of linear capabilities.
 % What is linear capabilities, split/splice
-A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
+A linear capability is a capability with no alias that cannot be duplicated.
+This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
 A capability machine with linear capabilities also needs special instructions to split and splice.
-Split and splice do exactly what their names indicate: Split splits a capability into two capabilities with the same permissions and with ranges that are disjointed but sum up to the range of the original capability, and splice does the inverse.
+Split and splice do exactly what their names suggest: Split splits a capability into two capabilities with the same permissions and with ranges that are disjointed and sum up to the range of the original capability, and splice does the inverse.
 
 % Stack capability linear
-The linear capabilities are put to use by using a linear capability for the stack pointer.
-When a trusted piece of code makes a call, it splits the stack capability in two, so it has a stack capability for its stack frame and one for the unused part of the stack.
-The stack capability for the caller stack frame is sealed and used as the data part of the sealed return pair.
-The stack capability for the remainder of the stack is given to the callee.
+The linear capabilities are put to use by requiring the stack capability to be linear.
+On call, the caller splits the stack capability in two, so they have a capability for their local stack frame and a capability for the unused part of the stack.
+The stack capability for their local stack frame is sealed and used as the data part of the sealed return pair, and the stack capability for the remainder of the stack is given to the callee.
 % Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
-As illustrated in Figure~\ref{fig:stack-ptr-abuse-prev}, this prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
-The trusted code knows that the stack capability is linear, so when it splits the stack capability, so it has a linear capability for its local stack frame, then it also knows that this part of memory cannot have an alias.
-The trusted code encapsulates the capability for its stack frame by sealing it which provides the guarantee that it can only be accessed when used with the appropriate code return pointer.
+Because the stack capability is linear the caller knows that the capability for their local stack frame cannot have an alias.
+This means that an adversary will need the stack capability the caller produced in order to access their local data.
+While the caller gives the capability to the adversary, it is only after the caller seals it which makes the capability opaque.
+This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
 
 % Prevention of attack 2
-In a traditional calling convention with a single stack, the stack is used as a call stack in the sense that it keeps track of the order of the calls.
-If everyone follows the calling convention, then a call pushes a stack frame to the call stack and a return pops the topmost stack frame from the call stack.
-This convention is not followed by the adversarial code in Figure~\ref{fig:ret-ptr-abuse} as the adversarial code skips two stack frames.
-In the presence of adversarial code, we need some enforcement mechanism that forces the adversarial code to follow the call stack order.
+In a traditional calling convention with a single stack, the stack serves as a call stack keeping track of the order calls where made in and thus in which order they should be returned to.
+A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
+However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
+This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
+In the presence of adversarial code, we need some enforcement mechanism that allows us to make sure that the order of the call stack is kept.
 One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
-The callee would have to give up this token on return to make sure that the token is only spent once.
-As it turns out, the stack capability for the unused part of the stack can be seen as such a token, so when the callee returns, we expect them to return the stack capability we gave them in the call.
-When we receive a stack capability on return, we will try to splice it with the now unsealed stack capability for our private stack frame.
-If this is successful, then we know that the two capabilities are adjacent.
-If it is unsuccessful or if no stack capability is returned, then we cannot determine whether our stack frame is the topmost.
-The issue illustrated in Figure~\ref{fig:ret-ptr-abuse} is prevented by splicing as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}.
+The callee would have to provide this token on return to prove that it is allowed to return to the caller, and on return the token would be taken back by the caller to prevent it from being spent multiple times.
+As it turns out, the stack capability for the unused part of the stack can be used as such a token in the following way:
+On return the callee has to give back the stack capability they were given on invocation.
+When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack.
+They do this by attempting to restore the stack capability from before the call by splicing the return token with the stack capability for the local stack frame which at this point has been unsealed again.
+If the splice is successful, then the caller knows that the two capabilities are adjacent.
+On the other hand, if the splice fails, then they are alerted to the fact that their stack frame may not be the topmost.
+StackTokens uses this approach, and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev} it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
 \begin{figure}
@@
 % Non-empty trusted stack frames.
-In this token-based scheme, it is important that the caller keeps a bit of the token which corresponds to having a non-empty stack frame.
-In a traditional C-like calling convention, this is natural as the caller needs to push information such as the old stack pointer and the old program counter to the stack.
-On a capability machine with CHERI's sealed capabilities, this is not necessary as the old program counter capability and the capability for the caller's stack frame is sealed in the return pair, and they will become unsealed when jumped to.
-If the caller has an empty stack frame, then they have entirely given up all of their return token which practically means that they are not present on the call stack and thus have no way verify whether it is their turn to be returned to.
+In order for a call to have a presence on the call stack, its stack frame must be non-empty.
+If we could have empty stack frames on the call stack, then it would be impossible to tell whether the topmost non-empty stack frame has an empty stack frame on top of it.
+Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in StackTokens these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
+This means that a caller using StackTokens needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
+There is also a more practical reason for a StackTokens caller to make sure their stack frame is non-empty: They need a bit of the stack capability in order to perform the splice that verifies the validity of the return token.
 
@@
 % + more
-We check that the returned stack capability is adjacent to the stack capability for our local stack frame, but this does not guarantee that it is the same token as we gave away in the call.
-Specifically, we have no idea whether the callee has kept part of the token for themselves and only given us back the part that is adjacent to our token.
-If we do not make sure that the entire token is returned, then it is possible for an adversary to break WBCF.
-This is illustrated in Figure~\ref{fig:stk-base-abuse}: Some trusted code has created a stack frame with their local data and called a piece of untrusted code giving them access to the remainder of the stack.
-This situation is illustrated in Figure~\ref{fig:stack-base-abuse-a}.
-Now the adversarial code keeps the stack capability for the part of the stack adjacent to the trusted code's stack frame.
-The adversary does not seal this part in a return pair, but simply stores it on the heap.
-At this point the adversary calls the trusted code with the remainder of the stack capability as illustrated in Figure~\ref{fig:stack-base-abuse-b}.
-At some point the trusted code calls the adversary again leaving a second stack frame on the stack an giving the adversary the rest.
-At this point, the adversary has access to the stack capability passed in the second call performed by the trusted code as well as the stack capability stored away after the first call (Figure~\ref{fig:stack-base-abuse-c}).
-Now the adversary has access to a stack capability for part of the stack adjacent to the first stack frame of the trusted caller.
-The adversary can use this stack capability to return from the first call without ever returning from the second call (Figure~\ref{fig:stack-base-abuse-d}).
+At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
+However, this does not make sure that the caller's stack frame is on top of the call stack.
+The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
+An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they later can return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
+This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
+The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
+The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
+At this point (Figure~\ref{fig:stack-base-abuse-b})), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
+For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
+In StackTokens, the caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to the base address of the return token they gave to the callee.
+In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
 
@@
 % The other attack.
+At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
+However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
+In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
+To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
+The trusted component executes using the stack capability given to them and at some point they invoke the callback creating return tokens and so on.
+The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
+Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback taking all the precautions we have described so far.
+At this point, the adversary has everything they need to return from either of the two callbacks.
+Specifically, they could return from the first callback invocation before they have returned from the second which would break WBCF.
+From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
+They are both correct as the adversarial component has created two call stacks, one for each of the components.
+In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
+We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
+A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
+This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
+
+Having a globally decided stack address also solves another challenge.
+Namely, the stack capability has to be linear in order for a caller to use it as a token, so a caller has to make sure that a capability is in fact linear when it is proposed as the stack capability by an untrusted source.
+This is, however, implicitly checked when the caller checks that the stack capability governs the global stack base address.
+The reason is that the stack capability needs to start execution as a linear capability which means that the capability that governs the global stack base address to begin with is linear.
+Due to the non-duplicability of linear capability and the fact that they cannot become non-linear during execution, the caller knows that if they have a capability that governs the global stack base address, then it must be linear.
+
+% Return pointers/return seals
+The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter and the data part is the stack capability for the local stack frame of the caller.
+Both of the capabilities in the pair are of course sealed with the same seal.
+% One seal per return point
+All call points need to be associated with a unique seal (a return seal) that is only used to seal the capabilities in the return pair for that particular call point.
+The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack which would break WBCF.
+For similar reasons, we cannot allow return seals to be used to seal closures.
+% Don't be stupid
+Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
+This goes for direct leaks, such as leaving a seal in a register or writing it to adversarial memory, as well as indirect leaks, where a capability for reading a return seal (either directly or indirectly) from memory is leaked.
+
+% Note about them vs us, we do nothing that they couldn't do
+In the description of the StackToken calling scheme, it has sometimes been in terms of us ``them vs us''.
+This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
+However, StackTokens is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
+This means that any component can ensure WBCF and LSE by employing StackTokens.
 
 % Summary of CC
+In the this section, we have described our proposed calling convention, StackTokens.
+In this paragraph we summarize it:
+\begin{description}
+\item[Check the base address of the stack capability before and after calls] 
+\item[Make sure that local stack frames are non-empty]
+\item[Create token and data return capability on call] split the stack capability in two to get a stack capability for you local stack frame and a stack capability for the unused part of the stack.
+The stack capability for your local stack frame is sealed and used for the data part of the return pair.
+\item[Create code return capability on call] Seal the old program pointer.
+\item[Reasonable use of seals] return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
+\end{description}
 
@@
 \item Sharing stack references accross component boundaries is supported
+\item Other notions of well-bracketedness (specifically one would be to allow different stacks)
 \end{itemize}

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
proofreading

@@
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
-Generally, properties that hold in the source language but not in the target language, need to be somehow enforced by the compiler.
+Generally, properties that hold in the source language but not in the target language need to be somehow enforced by the compiler.
 Two properties that hold in many high-level source languages, but not in the assembly languages they are compiled to, are well-bracketed control flow and encapsulation of local state.
@@
 Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
-At the assembly level, this is not so obvious, as invoked functions get direct access to return pointers, that they are supposed to jump to a single time, at the end of their execution, but there is no guarantee that untrusted assembly code respects this intended usage.
-Particularly, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned before.
+At the assembly level, this is not so obvious.
+Invoked functions get direct access to return pointers that they are supposed to jump to a single time at the end of their execution.
+There is, however, no guarantee that untrusted assembly code respects this intended usage.
+Particularly, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.
 
@@
 However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
-
+% KJAA: Tror der er et bedre ord en obvious. Tror ikke man ville sige''This property is obvious`', med mindre man mener at egenskaben i sig selv er enkel. Maaske "immeadiate"?
 To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
-The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
+The virtual memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
 More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
@@
 The concrete mechanics of object capabilities varies between different capability machines.
+% KJAA: This 'sealed' and 'common seal' business is unclear to me.
 On a recent capability machine called CHERI, for example, they take the form of pairs of capabilities sealed with a common seal that represent the code and data parts of the closure.
@@
 Also, in the presence of higher-order values (e.g.
-function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to, in order to provide it the right stack pointer upon invocation and it is not clear how this can be done efficiently in the presence of large amounts of components.
+function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
+It is not clear how this can be done efficiently in the presence of large amounts of components.
 Finally, this approach does not allow passing stack references between components.
@@
 A more scalable approach retains a single stack that is shared between components.
-Enforcing WBCF and LSE in this approach, requires a way to temporarily provide stack and return capabilities to an untrusted component and revoke them after it returns.
+Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component, and to revoke them after it returns.
 While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
 For example, CHERI offers a form of {\itshape local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
-\citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and a number of security checks and measures, the two properties can be guaranteed.
+\citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
+% KJAA: 'Boundary crossing' is new here.
 However, a problem with this approach is that revoking the local stack and return capabilities on every boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
@@
 Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
-These capabilities are enforced by the hardware to be non-duplicable.
+% KJAA: Perhaps make it clear by who and when/where?
+These capabilities are non-duplicable, enforced by the hardware.
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities, together with some security checks, then allows us to guarantee WBCF and LSE, without large overhead on boundary crossings and particularly no need for clearing large blocks of memory.
+The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings, particularly without the need for clearing large blocks of memory.
 
@@
 Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To accomodate this, we define a second operational semantics for our target language, which has a native well-bracketed stack and primitive ways to do calls and returns.
+To accomodate this, we define a second operational semantics for our target language, with a native well-bracketed stack and primitive ways to do calls and returns.
 This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
-As such, these components can be sure that they will only ever interact with well-behaved other components that respect our desirable properties.
-To express security of our calling convention, we then show that considering the same components in the original semantics, does not give adversaries additional ways to interact with them. 
+As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
+To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
 More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.
@@
 
-This approach expresses what it means to enforce our desirable properties, in a general way and makes it clear that we can support a very general class of programs.
+This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
 Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
@@
 Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
+% KJAA: Again with the sealed pair. Again, this idea is unclear.
 They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
@@
 %informally explain how an adversary may try to abuse stack and return caps
-While the above adds extra protection, it is important to understand that the above is not sufficient for enforcing WBCF and LSE.
+While the above adds extra protection, it is important to understand that it is not sufficient for enforcing WBCF and LSE.
 While adversaries only get access to capabilities that they are allowed to use during the stack frame, they have ways to store these capabilities and use them when they are no longer allowed to do so.
@@
 The adversary may have stored the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed to use it.
-The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call an thus breaks WBCF.
+The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.
 
@@
 The linear capabilities are put to use by requiring the stack capability to be linear.
-On call, the caller splits the stack capability in two, so they have a capability for their local stack frame and a capability for the unused part of the stack.
+On call, the caller splits the stack capability in two, such that they have a capability for their local stack frame and a capability for the unused part of the stack.
 The stack capability for their local stack frame is sealed and used as the data part of the sealed return pair, and the stack capability for the remainder of the stack is given to the callee.
@@
 % The other attack.
+% KJAA: It is unclear to me what 'at this point' refers to.
 At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
@@
 To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
-The trusted component executes using the stack capability given to them and at some point they invoke the callback creating return tokens and so on.
+The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
 The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
-Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback taking all the precautions we have described so far.
+Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
 At this point, the adversary has everything they need to return from either of the two callbacks.
-Specifically, they could return from the first callback invocation before they have returned from the second which would break WBCF.
+Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
 From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
@@
 % Return pointers/return seals
-The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter and the data part is the stack capability for the local stack frame of the caller.
+The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
 Both of the capabilities in the pair are of course sealed with the same seal.
@@
 % Note about them vs us, we do nothing that they couldn't do
-In the description of the StackToken calling scheme, it has sometimes been in terms of us ``them vs us''.
+In the description of the StackToken calling scheme, it has sometimes been in terms of ``them vs us''.
 This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
@@
 This means that any component can ensure WBCF and LSE by employing StackTokens.
+% KJAA: ^-- Assuming that there is a global stack address capability, no? Can I "suddenly" begin to impose StackTokens?
 

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
A few changes to the introduction based on Kris' feedback.

@@
 Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
-At the assembly level, this is not so obvious.
+At the assembly level, this property does not hold immediately. 
 Invoked functions get direct access to return pointers that they are supposed to jump to a single time at the end of their execution.
@@
 Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not have been read or modified when the invoked function returns.
-At the assembly level, this property is also far from obvious.
+At the assembly level, this property also does not hold immediately.
 The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
 However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
-% KJAA: Tror der er et bedre ord en obvious. Tror ikke man ville sige''This property is obvious`', med mindre man mener at egenskaben i sig selv er enkel. Maaske "immeadiate"?
 To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
-The virtual memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
+The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
 More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
@@
 % KJAA: This 'sealed' and 'common seal' business is unclear to me.
-On a recent capability machine called CHERI, for example, they take the form of pairs of capabilities sealed with a common seal that represent the code and data parts of the closure.
-The pair is opaque, but is transparently unsealed by the hardware upon invocation~\citep{watson_capability_2015}.
+On a recent capability machine called CHERI, for example, they take the form of pairs of capabilities that represent the code and data parts of the closure.
+Each of the two capabilities are sealed with a common seal which make them opaque.
+The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015}.
 

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
Added figures with syntax and op sem for target language as well as fig with syntax of source language.

@@
 
+\usepackage{placeins}
+
 \usepackage{todonotes}
@@
 \input{../preamble}
+% Paper specific redefinitions of commands
+\renewcommand{\MemSeg}{\shareddom{MemSeg}}
+\renewcommand{\Reg}{\shareddom{RegFile}}
+\renewcommand{\RegName}{\shareddom{RegName}}
+\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
+\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
+\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
+\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
+\newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
 
@@
 
+\section{A Capability Machine with Sealing and Linear Capabilities}
+\label{sec:cap-mach-w-seal-and-lin}
+\begin{itemize}
+\item present our \emph{target language} and its operational semantics (excerpts)
+\item mention roughly what components look like
+\end{itemize}
+\begin{figure}[htb]
+  \centering
+  \[
+  \arraycolsep=1.4pt
+  \begin{array}{rrcl l rrcl}
+   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
+    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
+    &\aendbnf & \in & \Addr \cup \{\infty \} & & &\sealendbnf & \in & \Seal \cup \{\infty \}\\
+    \scbnf \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
+    \cbnf \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
+    \rbnf \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
+    &\Reg & \defeq & \RegName \fun \Word & & &\Mem & \defeq & \Addr \fun \Word\\
+    &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
+    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
+  \end{array}
+\]
+\[
+  \arraycolsep=1.4pt
+\begin{array}{rcl}
+\multicolumn{3}{l}{    \arraycolsep=0pt
+      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
+}\\
+\Instr &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}}\mid\tgetp{\com{r}}{\com{r}} \mid\\
+      & &    \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
+  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\
+      & & \tsplice{\com{r}}{\com{r}}{\com{r}} 
+\end{array}
+\]
+  \caption{The syntax of our capbility machine with seals and linera capbilities.}
+  \label{fig:target-syntax}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \[
+    \Phi \step
+  \begin{cases}
+    \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) &
+    \arraycolsep=0pt
+      \begin{array}[t]{l}
+        \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \wedge
+        \baddr \le \aaddr \le \eaddr \wedge \perm \in \{\rwx,\rx\}
+      \end{array} \\
+      \failed & \totherwise
+  \end{cases}
+\]
+\[
+  \updPcAddr{\Phi} =
+  \begin{cases}
+    \Phi\updReg{\pcreg}{w} & \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge w = ((\perm,\lin),\baddr,\eaddr,\aaddr+1)\\
+    \Phi  & \totherwise
+  \end{cases}
+\]
+\[
+  \linCons{w} =
+  \begin{cases}
+    0 & \isLinear{w} \\
+    w & \totherwise
+  \end{cases}
+\] 
+\[
+  \xjmpres{c_1,c_2,\Phi} =
+  \begin{cases}
+    \Phi\updReg{\pcreg}{c_1}\updReg{\rdata}{c_2} & \nonExec{c_2} \\
+    \failed & \totherwise
+  \end{cases}
+\]
+  \begin{tabular}{|>{$}c<{$}|>{$}p{3.4cm}<{$}|>{\raggedright\arraybackslash}p{7cm}|}
+    \hline
+    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
+    \hline
+    \halt                                        & \halted & \\
+    \hline
+    \fail                                        & \failed & \\
+    \hline
+    \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{r}{\rn}} & $\rn \in \ints$\\
+    \hline
+    \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{\rn,r}{w_1,w_2}} & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\rn}$ \\
+    \hline
+    \load{r_1}{r_2}                              & \updPcAddr{\Phi\updReg{r_1}{w_1}\update{\mem.\aaddr}{w_\aaddr}} & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $w_a = \linCons{w_1}$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$\\
+    \hline
+    \store{r_1}{r_2}                             & \updPcAddr{\Phi\updReg{r_2}{w_2}\update{\mem.\aaddr}{\Phi(r_2)}} & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
+    \hline
+    \geta{r_1}{r_2}                              & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\_,\aaddr)$ or $\Phi(r_2) = \seal{\_,\_,\aaddr}$, then $w = \aaddr$ and otherwise $w = -1$\\
+    \hline
+    \jmp{r}    &\Phi\updReg{r,\pcreg}{w,\Phi(r)} & $w = \linCons{\Phi(r)}$\\
+    \hline
+    \xjmp{r_1}{r_2}                              & \Phi' & $\Phi(r_1) = \sealed{\sigma,c_1}$ and $\Phi(r_2) = \sealed{\sigma,c_2}$ and $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$ and $\Phi' = \xjmpres{c_1,c_2,\Phi\updReg{r_1,r_2}{w_1,w_2}}$  \\
+    \hline
+    \cca{r}{\rn}                                 &\updPcAddr{\Phi\updReg{r}{w}} & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
+    \hline
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{\Phi\updReg{r_3,r_1,r_2}{w,c_1,c_2}} & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_1)}$\\
+    \hline
+    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{\Phi\updReg{r_2,r_3,r_1}{w_2,w_3,c}}& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
+    \hline
+    \cseal{r_1}{r_2}                             & \Phi\updReg{r_1}{\vsc} & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
+    \hline
+    \multicolumn{3}{|c|}{\dots} \\
+    \hline
+    \_                                           & \failed & \totherwise \\
+    \hline
+  \end{tabular}
+\caption{An excerpt of the operational semantics of our capability machine with sealing and linear capabilities.}
+  \label{fig:target-op-sem}
+\end{figure}
+\FloatBarrier
 \section{Linear Stack and Return Capabilities}
-\todo[inline]{Make sealed capabilities more clear by greying out.}
+\todo[inline]{Make sealed capabilities more clear in figures.}
 \begin{figure}
@@
 
-\begin{itemize}
-\item informally explain how an adversary may try to abuse stack and return caps
-\item informally explain how we prevent this using linear capabilities
-\item use the tikz pictures from the PriSC presentation to explain all of this
-\end{itemize}
-
-\section{A Capability Machine with Sealing and Linear Capabilities}
-\label{sec:capab-mach-with}
-\begin{itemize}
-\item present our \emph{target language} and its operational semantics (excerpts)
-\item mention roughly what components look like
-\end{itemize}
+% \begin{itemize}
+% \item informally explain how an adversary may try to abuse stack and return caps
+% \item informally explain how we prevent this using linear capabilities
+% \item use the tikz pictures from the PriSC presentation to explain all of this
+% \end{itemize}
 
@@
 \label{sec:form-secur-with}
+\begin{figure}[htb]
+  \centering
+  \[
+    \arraycolsep=1.4pt
+    \begin{array}{rcl}
+      \src{\SealableCaps} & \defbnf& \SealableCaps \mid \src{\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}} \mid \src{\retptrd(\basebnf,\aendbnf)}\\
+                          & &  \mid \src{\retptrc(\basebnf,\aendbnf,\addrbnf)}\\
+      \multicolumn{3}{c}{
+      \begin{array}{lcrclcr}
+        \src{\StkFrame} & \defeq & \src{\Addr \times \MemSeg} & \phantom{skipskipsip} & \src{\Stack} & \defeq & \src{ \StkFrame^*}
+      \end{array}
+                                                                                                                }\\
+      \src{\ExecConf} & \defeq & \Mem \times \Reg \; \src{\times \; \Stack \times \MemSeg} \\
+    \end{array}
+  \] 
+\[
+  \begin{array}{rcl}
+    \offpc,\offsigma & \in & \nats \\
+    \src{\Instr} & \defbnf &  \Instr \mid \scall{\offpc,\offsigma}{r}{r}
+  \end{array}
+\]
+  \caption{The syntax of the capability machine with a built-in stack.}
+  \label{fig:src-syntax}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \label{fig:src-op-sem}
+  \caption{An excerpt of the operational semantics of the capability machine with a built-in stack.}
+\end{figure}
+
 \begin{itemize}
@@
 Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
-The capability machine formalized in Section~\ref{sec:capab-mach-with} is modeled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
+The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modeled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
 This is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
Fixed encoding issue
Added command for StackTokens name

@@
 %\copyrightyear{2018}           %% If different from \acmYear
+\usepackage[utf8]{inputenc}
 
@@
 
+\newcommand{\stktokens}{\textsc{StackTokens}}
+
 \begin{document}
@@
 %% Title information
-\title{StackTokens: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
+\title{\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
 % \titlenote{with title note}
@@
 \begin{abstract}
-  We propose and study StackTokens: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
+  We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
   The calling convention is based on linear capabilities, a type of capabilities that are prevented from being duplicated by the hardware.
@@
 
-In this work, we propose and study StackTokens: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
+In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
 Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
@@
 \FloatBarrier
+% \begin{itemize}
+% \item present our \emph{target language} and its operational semantics (excerpts)
+% \item mention roughly what components look like
+% \end{itemize}
 \section{Linear Stack and Return Capabilities}
@@
 
-Our StackTokens calling convention is based on a traditional single stack, shared between all components.
+Our \stktokens{} calling convention is based on a traditional single stack, shared between all components.
 As we are on a capability machine, it is natural to add some extra protection to stack and return pointers.
@@
 On the other hand, if the splice fails, then they are alerted to the fact that their stack frame may not be the topmost.
-StackTokens uses this approach, and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev} it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
+\stktokens{} uses this approach, and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev} it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
 \begin{figure}
@@
 If we could have empty stack frames on the call stack, then it would be impossible to tell whether the topmost non-empty stack frame has an empty stack frame on top of it.
-Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in StackTokens these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
-This means that a caller using StackTokens needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
-There is also a more practical reason for a StackTokens caller to make sure their stack frame is non-empty: They need a bit of the stack capability in order to perform the splice that verifies the validity of the return token.
+Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in \stktokens{} these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
+This means that a caller using \stktokens{} needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
+There is also a more practical reason for a \stktokens{} caller to make sure their stack frame is non-empty: They need a bit of the stack capability in order to perform the splice that verifies the validity of the return token.
 
@@
 For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In StackTokens, the caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to the base address of the return token they gave to the callee.
+In \stktokens{}, the caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to the base address of the return token they gave to the callee.
 In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
@@
 This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
-However, StackTokens is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
-This means that any component can ensure WBCF and LSE by employing StackTokens.
-% KJAA: ^-- Assuming that there is a global stack address capability, no? Can I "suddenly" begin to impose StackTokens?
+However, \stktokens{} is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
+This means that any component can ensure WBCF and LSE by employing \stktokens{}.
+% KJAA: ^-- Assuming that there is a global stack address capability, no? Can I "suddenly" begin to impose \stktokens{}?
 
 % Summary of CC
-In the this section, we have described our proposed calling convention, StackTokens.
+In the this section, we have described our proposed calling convention, \stktokens{}.
 In this paragraph we summarize it:
@@
 This is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
-For simplicity, we have omitted features of CHERI that were not needed for modelling StackTokens (e.g.\ local capabilities, virtual memory).
+For simplicity, we have omitted features of CHERI that were not needed for modelling \stktokens{} (e.g.\ local capabilities, virtual memory).
 
@@
 This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
-Contrary to StackTokens, CFI can be implemented on commodity hardware rather than capability machines.
+Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
 However, its attacker model is different and its security goals are weaker.

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
Formatting of figure
Added outline for section about target machine.
Some work on target machine section.

@@
 \renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
+\renewcommand{\perm}{\var{p}}
+\renewcommand{\SealableCaps}{\shareddom{Sealables}}
+\renewcommand{\Cap}{\shareddom{NatTok}}
+
 \newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
@@
 \newcommand{\stktokens}{\textsc{StackTokens}}
+
+\newcommand{\trgcm}{\textsc{LCM}}
+\newcommand{\srccm}{\textsc{SLCM}}
 
@@
 \label{sec:cap-mach-w-seal-and-lin}
-\begin{itemize}
-\item present our \emph{target language} and its operational semantics (excerpts)
-\item mention roughly what components look like
-\end{itemize}
-\begin{figure}[htb]
+\begin{figure}[tb]
   \centering
@@
    \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
+    \aendbnf \in & \Addr \cup \{\infty \} & & & & \sealendbnf \in & \Seal \cup \{\infty \}& &\\
     \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
-    &\aendbnf & \in & \Addr \cup \{\infty \} & & &\sealendbnf & \in & \Seal \cup \{\infty \}\\
-    \scbnf \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
-    \cbnf \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
-    \rbnf \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
+    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
+    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
+    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
     &\Reg & \defeq & \RegName \fun \Word & & &\Mem & \defeq & \Addr \fun \Word\\
@@
 \end{figure}
+% Brief intro to section and this cap machine:
+% Cap machine inspired by CHERI, but with linear capabilities
+In this section, we introduce a simple but representative capability machine with linear capabilities.
+We will later introduce a capability machine similar to the one in this section, so to easily distinguish the two, we will call this one \trgcm{} (Linear Capability Machine).
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main exception.
+In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address and unbounded integers.
+\lau{Do we have a good reason to assume an infinite address space for this work?
+  We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
+
+\begin{wrapfigure}{r}{0.33\linewidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (rwx) {$\rwx$};
+    \node[main node] (rx) [below right of=rwx] {$\rx$};
+
+    \node[main node] (rw) [below left of=rwx,] {$\rw$};
+    \node[main node] (r) [below right of=rw] {$\readonly$};
+    \node[main node] (0) [below of=r] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+
+    (rw) edge (r)
+    (r) edge (0)
+
+    (rwx) edge (rx)
+
+    (rw) edge (rwx)
+    (r) edge (rx);
+  \end{tikzpicture}
+
+  \caption{Permission hierarchy}
+  \label{fig:perm-hier}
+\end{wrapfigure}
+%%% Capabilities
+The concept of a capability is the cornerstone of any capability machine.
+In its simplest form, a capability is a permission and a range of authority.
+The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
+% Permissions and Range of authority
+The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}) where $\var{perm}$ is the permission and $[\var{base},\var{end}]$ is the range of authority.
+\trgcm{} only has memory capabilities, and we limit them to a number of sensible permissions making the available permissions: read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$).
+We define the preorder $\le$ on permissions as illustrated in Figure~\ref{fig:perm-hier}.
+% Pointer and linearity
+In addition to the two basic things, the capabilities also have a current address $\addr$ and a linearity $\lin$.
+The linearity of a capability can either be $\linear$ or $\normal$ where $\normal$ corresponds to a traditional capability.
+We will describe linear capabilities in detail when we get to the operational semantics of \trgcm{}.
+
+%%% Seals
+Any reasonable capability machine needs a way to setup boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
+%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
+On \trgcm{} we have sealed capabilities to achieve this.
+Syntactically, sealed capabilities add two things to the capability machine: Sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ and sealed capabilities $\sealed{\sigma,\vsc}$.
+% seals
+A set of seals consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$ that is represented by a natural number.
+% sealed caps
+A sealed capability consists of a sealable $\vsc$ and the seal $\sigma$ it has been sealed with.
+The sealables are sets of seals and capabilities, and we refer to the set of them as $\SealableCaps$.
+The system native tokens on \trgcm{} are the sealables and the sealed capabilities and we refer to this set as $\Cap$.
+
+%%% Domains
+The words on \trgcm{} are the system native tokens in $\Cap$ and data which is represented by $\ints$.
+We assume a finite set $\RegName$ of register names that at least contains $\{\pcreg , \rretd , \rretc , \rstk , \rdata , \rtmp{1} , \rtmp{2}\}$, and we define register files as a function from register names to words.
+A memory is a full memory and a memory segment is part of a memory.
+Both are defined as functions from addresses, defined as the natural numbers, to words, but the memory segment functions only need to be partial.
+$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution with an attempt to break the capability machine guarantees.
+An executable configuration is a register file and memory pair.
+
+% syntax (instructions)
+% opsem
+% move w/ lin
+% load w/ lin
+% cseal
+% xjmp
+% split/splice
+
+% Components ...
+% write xor exec
+% contents
+% linking
 
@@
   \[
+    \begin{array}{l}
     \Phi \step
@@
       \failed & \totherwise
-  \end{cases}
-\]
-\[
+  \end{cases}\\
   \updPcAddr{\Phi} =
@@
     \Phi  & \totherwise
-  \end{cases}
-\]
-\[
+  \end{cases}\\
   \linCons{w} =
@@
     w & \totherwise
-  \end{cases}
-\] 
-\[
+  \end{cases}\\
   \xjmpres{c_1,c_2,\Phi} =
@@
   \end{cases}
+  \end{array}
 \]

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
wording

@@
 % KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
+However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
 

### 2018-07-04 fully-abstract-comp/technical-report.tex
Typos and minor improvements to tech report.
Renamed tech report and moved stktokens macro to preamble file.

@@
 
-\title{Technical Report}
-
+\title{\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities\\
+  Technical Report with Proves and Details}
 \begin{document}
@@
 We will make the technical report up to date as soon as possible.
+Further, the use of coloring in the later parts of this document has not always been done with as much care as it ought to.
+This means that not all source specific things are colored blue, but it should be evident from the context that they belong to the source language.
 
@@
     \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}\\ 
-    &\rbnf & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
+    \rbnf \in & \RegName & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
     &\Reg & \defeq & \RegName \fun \Word\\
@@
   \begin{array}{rcl}
-    o & \in & \nats \\
+    \offpc,\offsigma & \in & \nats \\
     \src{i} & \defbnf &  \com{i} \mid \scall{\offpc,\offsigma}{r}{r}
@@
     \begin{array}[t]{l}
-      c_1 \neq \retptrc(\_) \tand \\
-      c_2 \neq \retptrd(\_) \tand \\
+      \src{c_1 \neq \retptrc(\_) \tand} \\
+      \src{c_2 \neq \retptrd(\_) \tand} \\
       \nonExec{c_2}

### 2018-07-04 fully-abstract-comp/technical-report.tex
Note on instruction set in tech report.

@@
 % At this juncture, it does, however, not seem necessary to have opaque because sealed capabilities can do their job.
+\subsection{Notes on linear capabilities}
+It seems reasonable to have enough instructions to let any program be able to make sufficient checks that it can verify that its execution won't fail.
+With our current instruction set, a load may fail if a linear capability happens to be located at a memory address that on attempts to load from with a capability without write permission.
+To make up for this, one could make an instruction that checks the linearity of a capability in memory without loading it.
+It may not be practical to make such an instruction if linearity is kept track as a field on each capability, but it may be tractable if linearity tags are kept track of in a table.
 

### 2018-07-04 fully-abstract-comp/popl2019/paper.tex
Renamed tech report and moved stktokens macro to preamble file.
More work on target machine section.

@@
 \newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
-
-\newcommand{\stktokens}{\textsc{StackTokens}}
 
@@
 }\\
-\Instr &\defbnf & \tfail \mid \thalt \mid \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}}\mid\tgetp{\com{r}}{\com{r}} \mid\\
-      & &    \tgetlin{\com{r}}{\com{r}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tstore{\com{r}}{\com{r}} \mid \tload{\com{r}}{\com{r}} \mid \tcca{\com{r}}{\com{\rn}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \\
-  & & \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tsetatob{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid \tcseal{\com{r}}{\com{r}} \mid \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\
-      & & \tsplice{\com{r}}{\com{r}}{\com{r}} 
+  \Instr &\defbnf & \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tload{\com{r}}{\com{r}} \mid \tstore{\com{r}}{\com{r}} \mid \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid\\
+         & & \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid\tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}}  \mid \\
+  & & \tcca{\com{r}}{\com{n\rn}} \mid \tsetatob{\com{r}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tcseal{\com{r}}{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid  \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\
+      & & \tsplice{\com{r}}{\com{r}}{\com{r}} \mid \tfail \mid \thalt 
 \end{array}
@@
   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
+
+%%% Capabilities
+The concept of a capability is the cornerstone of any capability machine.
+In its simplest form, a capability is a permission and a range of authority.
+The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
+% Permissions and Range of authority
+The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}) where $\var{perm}$ is the permission and $[\var{base},\var{end}]$ is the range of authority.
+\trgcm{} only has memory capabilities, and we limit them to a number of sensible permissions making the available permissions: read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$).
+We define the preorder $\le$ on permissions as illustrated in Figure~\ref{fig:perm-hier}.
+% Pointer and linearity
+In addition to the two basic things, the capabilities also have a current address $\addr$ and a linearity $\lin$.
+The linearity of a capability can either be $\linear$ or $\normal$ where $\normal$ corresponds to a traditional capability.
+% What is linear capabilities, split/splice
+A linear capability is a capability with no alias that cannot be duplicated.
+This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
+We will see this later in this section when we get to the operational semantics.
 
@@
 \end{wrapfigure}
-%%% Capabilities
-The concept of a capability is the cornerstone of any capability machine.
-In its simplest form, a capability is a permission and a range of authority.
-The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
-% Permissions and Range of authority
-The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}) where $\var{perm}$ is the permission and $[\var{base},\var{end}]$ is the range of authority.
-\trgcm{} only has memory capabilities, and we limit them to a number of sensible permissions making the available permissions: read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$).
-We define the preorder $\le$ on permissions as illustrated in Figure~\ref{fig:perm-hier}.
-% Pointer and linearity
-In addition to the two basic things, the capabilities also have a current address $\addr$ and a linearity $\lin$.
-The linearity of a capability can either be $\linear$ or $\normal$ where $\normal$ corresponds to a traditional capability.
-We will describe linear capabilities in detail when we get to the operational semantics of \trgcm{}.
 
@@
 The words on \trgcm{} are the system native tokens in $\Cap$ and data which is represented by $\ints$.
-We assume a finite set $\RegName$ of register names that at least contains $\{\pcreg , \rretd , \rretc , \rstk , \rdata , \rtmp{1} , \rtmp{2}\}$, and we define register files as a function from register names to words.
+We assume a finite set $\RegName$ of register names that at least contains registers $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as a function from register names to words.
 A memory is a full memory and a memory segment is part of a memory.
@@
 
-% syntax (instructions)
-% opsem
-% move w/ lin
-% load w/ lin
-% cseal
-% xjmp
-% split/splice
-
-% Components ...
-% write xor exec
-% contents
-% linking
-
-\begin{figure}[h]
+\begin{figure}[p]
   \centering
@@
 \end{figure}
-\FloatBarrier
+% syntax (instructions)
+\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability machine specific instructions.
+The instructions that can be considered standard are: unconditional and a conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
+The simplest of the capability machine specific instructions simply inspect system native tokens or project one of their fields.
+These are \texttt{gettype} (for getting the type of a system token) and \texttt{getp}, \texttt{getl}, \texttt{getb}, \texttt{gete}, and \texttt{geta} (for projecting permission, linearity, base, end, and current address(/seal), respectively).
+There are two instructions for manipulating the current address(/seal) of a capability(/set of seals): \texttt{cca} changes the current address(/seal) of a capability(/set of seals) by some offset, and \texttt{seta2b} set the current address(/seal) of a capability(/set of seals) to the base address(/seal).
+The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
+Generally speaking, a capability machine needs an instruction for reducing the range of authority of an capability.
+Because \trgcm{} has linear capabilities the instruction for this splits the capability in two rather than reducing the range of authority (instruction \texttt{split}).
+\trgcm{} also has an instruction that does the inverse of this, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
+\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities (\texttt{xjmp}).
+Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).
+
+An representative excerpt of the operational semantics is shown in Figure~\ref{fig:target-op-sem}.
+% opsem
+% move w/ lin
+% load w/ lin
+% cseal
+% xjmp
+% split/splice
+\todo[inline]{Write this.}
+
+% Components ...
+% write xor exec
+% contents
+% linking
+\todo[inline]{Write this.}
+
+Some simplifications has been made in this presentation of \trgcm{}. We refer to the accompanying technical report\citep{technical_report} for all details.
+
+%\FloatBarrier
 % \begin{itemize}
@@
 The two properties can be enforced with a calling convention that utilizes the guarantees of linear capabilities.
-% What is linear capabilities, split/splice
-A linear capability is a capability with no alias that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
-A capability machine with linear capabilities also needs special instructions to split and splice.
-Split and splice do exactly what their names suggest: Split splits a capability into two capabilities with the same permissions and with ranges that are disjointed and sum up to the range of the original capability, and splice does the inverse.
 

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
Adapt a figure explaining the full abstraction proof from a previous paper.

@@
 
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
+    % \draw[help lines,yellow] (0,0) grid (10,7);
+    \node at (5,4.7) { ${\src{C_1}\mathrel{\sconeq} \src{C_2}}$ };
+
+    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{C_1}} \sterm[]{\gc}{}}$ };
+    \node at (5,4) { $\mathrel{\Ra}$ };
+    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{C_2}} \sterm[]{\gc}{}}$ };
+
+    \node at (4.35,2.8) { (1) };
+    \node at (5,3.6) { (2) };
+    \node at (5.65,2.8) { (3) };
+
+    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
+
+    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
+
+    \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{C_2}} \cong \comp{{\src{C_2}}}$};
+    \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{C_1}} \cong \comp{{\src{C_1}}}$};
+    % \node at (9,2.7) { $e  {\src{C_1}} \cong \comp{{\src{C_1}}} : tau$ };
+    % \node at (8.7,3.3) { $ {\comp{\trg{\context}}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
+    % \node at (.8,3) { $e  {\src{C_1}} \cong \comp{{\src{C_1}}} : tau$ };
+
+    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{C_1}} \term[]{}}$ };
+    \node at (5,2.1) { $\overset{?}{\Ra}$ };
+    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{C_2}} \term[]{}}$ };
+
+    \node at (5,1.3) { ${\src{C_1}}\mathrel{\overset{?}{\tconeq}}{\src{C_2}}$ };
+
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
+  \end{tikzpicture}
+  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
+\end{figure}
+
 \section{Discussion}

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
Fix latex issues.
inserted missing variables in target machine syntax figure
Reordered trg machine op sem figure and fixed typo.
First draft of section about target machine done.

--- 2018-07-04 fully-abstract-comp/popl2019/paper.tex
@@
     r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
-    &\Reg & \defeq & \RegName \fun \Word & & &\Mem & \defeq & \Addr \fun \Word\\
+    \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
     &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
@@
     \fail                                        & \failed & \\
-    \hline
-    \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{r}{\rn}} & $\rn \in \ints$\\
+    % \hline
+    % \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{r}{\rn}} & $\rn \in \ints$\\
     \hline
@@
     \hline
+    \cca{r}{\rn}                                 &\updPcAddr{\Phi\updReg{r}{w}} & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
+    \hline
     \jmp{r}    &\Phi\updReg{r,\pcreg}{w,\Phi(r)} & $w = \linCons{\Phi(r)}$\\
@@
     \xjmp{r_1}{r_2}                              & \Phi' & $\Phi(r_1) = \sealed{\sigma,c_1}$ and $\Phi(r_2) = \sealed{\sigma,c_2}$ and $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$ and $\Phi' = \xjmpres{c_1,c_2,\Phi\updReg{r_1,r_2}{w_1,w_2}}$  \\
-    \hline
-    \cca{r}{\rn}                                 &\updPcAddr{\Phi\updReg{r}{w}} & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
     \hline
@@
     \hline
-    \cseal{r_1}{r_2}                             & \Phi\updReg{r_1}{\vsc} & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
+    \cseal{r_1}{r_2}                             & \updPcAddr{\Phi\updReg{r_1}{\vsc}} & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
     \hline
@@
 
-An representative excerpt of the operational semantics is shown in Figure~\ref{fig:target-op-sem}.
 % opsem
+The operational semantics of \trgcm{} is displayed in Figure~\ref{fig:target-op-sem}.
+The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
+The executed instruction is determined by the capability in $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
+In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the range of addresses where the capability can operate.
+If both things are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relatively to $\Phi$.
+The interpretation of some of the instructions are also displayed in Figure~\ref{fig:target-op-sem}.
+In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
+The instructions that stop execution or change the flow of execution do not use $\updPcAddr{}$.
+For instance, the \texttt{halt} and \texttt{fail} instructions are simply interpreted as the $\halted$ and $\failed$ configurations, respectively, and they do not use $\updPcAddr{}$.
+
 % move w/ lin
-% load w/ lin
+The \texttt{move} instruction simply moves a word from one register to another.
+It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
+This means that when a linear capability is moved, the source register should be cleared.
+To take care of this and prevent duplicating the operational semantics, we use the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
+When the \texttt{move} instruction is performed, the source register $r_2$ is overwritten with $\linCons{\Phi(r_2)}$ which clears the register if the contents was linear and leaves it unchanged otherwise.
+
+% load w/ lin (and store)
+The \texttt{store} and \texttt{load} instructions are fairly standard: They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority, and they clear the source of a linear capability transfer.
+Linear capabilities introduce one complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
+Clearing a memory address corresponds to writing $0$ to the memory, but in order to write to memory we need to provide a capability with write-permission.
+This means that to load a linear capability from memory, we have to use a capability with read and write permission.
+
+% geta, cca
+The instructions \texttt{geta} and \texttt{cca} are fairly simple: \texttt{geta} projects the current address or current seal from a capability or a set of seals, respectively, and it returns the error value $-1$ for data and sealed capabilities.
+The \texttt{cca} instruction changes the current address or seal of a capability or set of seals, respectively, by a given offset.
+By making an in-place update, the instruction does not need to take special care in order to maintain linearity.
+The \texttt{jmp} instruction is a fairly simple jump that just sets the $\pcreg$-register.
+
+%% Sealing
+The operational side of the sealing in \trgcm{} consists of two instructions: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
 % cseal
+Given a sealable $\vsc$ and a set of seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
 % xjmp
+Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities unseals them and put one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the capability placed in $\rdata$) is non-executable.
+A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment of the closure.
+Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
+Some care should be taken when sealing as reuse of the same seal for multiple closures would make it possible to jump to the code of one capability with the environment of another.
+% xjmp to unseal
+\trgcm{} does not have an instruction for unsealing single sealed capabilities directly, but most sealed capabilities can be achieved unsealed using \texttt{xjmp} and a set of seals with the seal used to seal the sealed capability.
+
 % split/splice
-\todo[inline]{Write this.}
+Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
+For normal capabilities, reduction of authority can be done without actually giving up any authority if the capability is duplicated first.
+With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
+In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
+The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as it range of authority.
+Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
+With \texttt{split} we can reduce the range of authority of a linear capability without losing any authority as we retain it in the other capability.
+However, if we only have \texttt{split}, then in some sense splitting a capability is not completely lossless as we may have authority over the same range of addresses, but we have that authority in two capabilities rather than just one.
+In order to make up for this, we have a \texttt{splice} instruction that essentially does the inverse of \texttt{split}.
+Namely given two capabilities with adjacent ranges of authority and the same permissions and linearity, it splices them together into one capability.
+The two instructions work in the same way for seal sets.
+We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
+This would also allow linear capabilities to have aliases, but only by linear capabilities with disjointed permissions.
 
-% Components ...
+At this time, the executable configuration is our only way to specify the state of the capability machine.
+We would, however, much prefer a notion of a program that only describe the essential parts of an entity on the machine and only later instrument the program with whatever necessary to make it into an executable configuration.
+In later sections, we describe different entities on the capability machine that interact with each other.
+In order for them to interact, we need to be able to link them together and to this end, we introduce a notion of a component.
+A component is basically a program with open ends in the form of imports.
+A component also have exports allowing other programs to link with them.
+We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.
+% Structure
+A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
+The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
+The exports are words associated with a symbol.
+%base = library
+A base component can be seen as library implementations because it does not specify a main entry point of execution.
+Components are either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
+% Linking .
+Components are combined into new components by linking them together.
+Libraries can be linked together and incomplete programs can be linked with libraries, but two incomplete programs cannot be linked with each other.
+Two components can be linked when their memories, seals, and linear addresses disjointed.
+They are combined by taking the union of each of their constituents.
+For every import that is satisfied by an export of the other component, in the sense that they have the same symbol, the data memory is updated to have the exported word on the imported address.
+The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
+Linking is defined in Figure~\ref{fig:target-component-and-linking}.
+
+% Programs
+We can now define the notion of a program as well as a context.
+\begin{definition}[Programs and Contexts]
+  We define a \emph{program} to be a component $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
+
+  A \emph{context} for a component $\var{comp}$ is another component $\var{comp}'$ such that $\var{comp} \bowtie \var{comp}'$ is a program.
+\end{definition}
+% Getting an executable from a component - postpone this to later, but mention:
 % write xor exec
-% contents
-% linking
-\todo[inline]{Write this.}
+In Section~\ref{sec:form-secur-with}, we will get back to what a program is instrumented with in order to create an executable configuration. 
+We will, however, at this point mention that all programs will be instrumented with the Write-XOR-Execute,\todo{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+%Further, we will only consider a specific class of well-formed components, but we will get back to this.
 
-Some simplifications has been made in this presentation of \trgcm{}. We refer to the accompanying technical report\citep{technical_report} for all details.
+\begin{figure}[htb]
+  \centering
+  \[
+    \begin{array}{rl c rl c rl}
+    s &\in \Symbol & \phantom{make} &
+    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
+    \mathrm{export} &\mathrel{::=} s \mapsto w\\
+    \var{comp}_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
+    \var{comp} & \multicolumn{7}{l}{\mathrel{::=} \var{comp}_0 \mid  (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
+  \end{array}
+\]
+\begin{mathpar}
+  \inference{
+    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
+    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
+    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
+    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
+    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
+    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
+    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
+    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
+    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
+    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
+    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
+  } {
+    \var{comp}_3 = \var{comp}_1 \bowtie \var{comp}_2
+  }
+  \and
+  \inference{
+    \var{comp}_0'' = \var{comp}_0 \bowtie \var{comp}_0'
+  }{
+    (\var{comp}_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \var{comp}_0 \bowtie (\var{comp}_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \var{comp}_0'
+  }
+\end{mathpar}
+% \begin{definition}[Initial execution configuration]
+%   \begin{mathpar}
+%     \inference{
+%       c_{\mathrm{main},c} = \sealed{\sigma_1, c_{\mathrm{main},c}'} &
+%       c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'} &
+%       \sigma_1 = \sigma_2\\
+%       \nonExec{c_{\mathrm{main},d}'}&
+%       \reg(\pcreg) = c_{\mathrm{main},c}' & 
+%       \reg(\rdata) = c_{\mathrm{main},d}' \\
+%       \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
+%       \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
+%       \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0\\
+%       \range{\ms_\stk} = \{0\}&
+%       \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
+%       [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
+%       \overline{\var{import}} = \emptyset
+%     }{
+%       ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
+%     }
+%   \end{mathpar}
+% \end{definition}
+% \begin{definition}[Plugging a program into a context]
+%   When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
+%   then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
+% \end{definition}
+  \caption{Components and linking of components}
+  \label{fig:target-component-and-linking}
+\end{figure}
+
+Some simplifications has been made in this presentation of \trgcm{} in this section. We refer to the accompanying technical report\citep{technical_report} for all details.
 

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
note for source machine section.

@@
 \item present our source language, its operational semantics (excerpts)
+\item tell more about components, specifically well-formed according to judgement
 \item mention our assumption of reasonability
@@
 \end{itemize}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
+    % \draw[help lines,yellow] (0,0) grid (10,7);
+    \node at (5,4.7) { ${\src{C_1}\mathrel{\sconeq} \src{C_2}}$ };
+
+    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{C_1}} \sterm[]{\gc}{}}$ };
+    \node at (5,4) { $\mathrel{\Rightarrow}$ };
+    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{C_2}} \sterm[]{\gc}{}}$ };
+
+    \node at (4.35,2.8) { (1) };
+    \node at (5,3.6) { (2) };
+    \node at (5.65,2.8) { (3) };
+
+    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
+
+    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
+
+    \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{C_2}} \cong \comp{{\src{C_2}}}$};
+    \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{C_1}} \cong \comp{{\src{C_1}}}$};
+    % \node at (9,2.7) { $e  {\src{C_1}} \cong \comp{{\src{C_1}}} : tau$ };
+    % \node at (8.7,3.3) { $ {\comp{\trg{\context}}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
+    % \node at (.8,3) { $e  {\src{C_1}} \cong \comp{{\src{C_1}}} : tau$ };
+
+    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{C_1}} \term[]{}}$ };
+    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
+    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{C_2}} \term[]{}}$ };
+
+    \node at (5,1.3) { ${\src{C_1}}\mathrel{\overset{?}{\tconeq}}{\src{C_2}}$ };
+
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
+  \end{tikzpicture}
+  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
+\end{figure}
 

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
minor changes to trg machine section.

@@
 The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
-Linking is defined in Figure~\ref{fig:target-component-and-linking}.
+In Figure~\ref{fig:target-component-and-linking}, $\bowtie$ defines linking.
 
@@
 \begin{definition}[Programs and Contexts]
+  \label{def:program-and-context}
   We define a \emph{program} to be a component $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
@@
 % write xor exec
-In Section~\ref{sec:form-secur-with}, we will get back to what a program is instrumented with in order to create an executable configuration. 
-We will, however, at this point mention that all programs will be instrumented with the Write-XOR-Execute,\todo{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+In Section~\ref{sec:form-secur-with}, we will get back to what a program is instrumented with in order to create an executable configuration. \todo[inline]{Make sure we got back to this!}
+We will, however, at this point mention that all programs will be instrumented with the Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
 
+Some simplifications has been made in this presentation of \trgcm{} in this section. We refer to the accompanying technical report\citep{technical_report} for all details.
 \begin{figure}[htb]
@@
 
-Some simplifications has been made in this presentation of \trgcm{} in this section. We refer to the accompanying technical report\citep{technical_report} for all details.
-
 %\FloatBarrier

### 2018-07-05 fully-abstract-comp/technical-report.tex
Small variable name correction
Added a todo in tech report.
Remark about missing gc in source instruction interpretation.

@@
 \end{align*}
+On the source machine, the instruction interpretation also takes $\gc$, the global constants.
+It does, however, just pass it around and never makes changes to it, so we will leave it implicit.
+
 \dominique{Note: The above does not allow executing code that is on the stack in the source machine, while this is in principle allowed in the target.
@@
 \end{align*}
-
+\lau{(important) TODO 05-07-2018: It seems that there may be an issue related to splicing stk-ptr with a linear capability.
+  This would be okay on trg machine, but fail on src machine.
+  It may possibly be fixed simply by padding the stack with un-addressable memory (like we did with code memory).}
 Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
@@
   \var{export} &\mathrel{::=} s \mapsto w\\
-  \var{comp}_0 &\mathrel{::=} (\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)\\
+  \var{comp}_0 &\mathrel{::=} (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
   \var{comp} &\mathrel{::=} \var{comp}_0\\
@@
   \inference{
-    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
-    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
-    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
+    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
+    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
+    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
     \mscode[3] = \mscode[1] \uplus \mscode[2] \\

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
New name for the source machine: oLCM for overlay LCM.
Todo about a figure
short introduction to section about stktokens.

@@
 \newcommand{\trgcm}{\textsc{LCM}}
-\newcommand{\srccm}{\textsc{SLCM}}
+\newcommand{\srccm}{\textsc{oLCM}}
 
@@
 In Figure~\ref{fig:target-component-and-linking}, $\bowtie$ defines linking.
+\todo[inline]{Add better separation between component and linking in figure.}
 
@@
 \todo[inline]{Make sealed capabilities more clear in figures.}
+In this section, we introduce our calling convention \stktokens{} that insures LSE and WBCF.
+We will gradually explain each of the security measures \stktokens{} take and motivate them with the attacks they prevent.
+
 \begin{figure}

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
Minor textual changes to StkTokens section.

@@
 
-Our \stktokens{} calling convention is based on a traditional single stack, shared between all components.
+\stktokens{} is based on a traditional single stack, shared between all components.
 As we are on a capability machine, it is natural to add some extra protection to stack and return pointers.
@@
 %informally explain how an adversary may try to abuse stack and return caps
-While the above adds extra protection, it is important to understand that it is not sufficient for enforcing WBCF and LSE.
-While adversaries only get access to capabilities that they are allowed to use during the stack frame, they have ways to store these capabilities and use them when they are no longer allowed to do so.
+While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
+The caller provides the adversary with a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
@@
 As the examples illustrate, the standard memory and object-like capabilities (sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
-The two properties can be enforced with a calling convention that utilizes the guarantees of linear capabilities.
+The two properties can, however, be enforced with a calling convention that utilizes the guarantees of linear capabilities.
 

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
Introduction to FAOS section.

@@
 \label{sec:form-secur-with}
+A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
+Our proof uses a novel approach called full abstract overlay semantics.
+In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
+An overlay semantics is basically a different view on an existing semantics.
+For instance, it may have new syntactic constructs and have special semantics for these things, but they will correspond to something on the underlying machine.
+In the overlay semantics we are about to define for \trgcm{}, there is an explicit stack pointer construct on the machine, but on the underlying semantics the stack pointer corresponds directly to a linear capability for the part of memory the overlay semantics has deemed to be the stack.
+When we have defined the overlay semantics, we prove full-abstractness for a compilation from the overlay semantics to the underlying semantics. 
+Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as it is really just a different view on the same thing. 
+In this case, the compilation even turns out to be the identity function!
+\todo[inline]{Maybe also introduce reasonability here}
+
 \begin{figure}[htb]

### 2018-07-05 fully-abstract-comp/popl2019/paper.tex
Outline for FAOS section.

@@
 
+The overlay semantics we define for \trgcm{} (\srccm{}) is defined in Figure~\ref{fig:src-syntax} and \ref{fig:src-op-sem}.
+The main things it adds to \trgcm{} is a built-in stack that can be used to store local state as well as a \texttt{call} instruction and a new interpretation of \texttt{xjmp} that corresponds to returning.
+% Describe the things added to the syntax
+% Built-in stack
+% Stack pointer
+% Return pointers
+
 \begin{figure}[htb]
@@
 
-\begin{itemize}
-\item present our source language, its operational semantics (excerpts)
-\item tell more about components, specifically well-formed according to judgement
-\item mention our assumption of reasonability
-\item present the full abstraction theorem.
-\end{itemize}
+% Describe the changes to op sem
+% Step relation change, call case and global constants
+% Call semantics
+% New xjmp semantics (can be defined in terms of a new xjmpresult function!)
+% Example of alternative interpretation for the new syntactic constructs
+
+% Well-formed components
+
+% Reasonability
+
+% Full abstraction thm
+
+% \begin{itemize}
+% \item present our source language, its operational semantics (excerpts)
+% \item tell more about components, specifically well-formed according to judgement
+% \item mention our assumption of reasonability
+% \item present the full abstraction theorem.
+% \end{itemize}
 

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Reformulation.

@@
 % KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
+However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.
 

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Some minor rephrasings.

@@
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they have been mentioned in some technical documents.}
 % KJAA: Perhaps make it clear by who and when/where?
@@
 \trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main exception.
-In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address and unbounded integers.
+In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address space and unbounded integers.
 \lau{Do we have a good reason to assume an infinite address space for this work?
@@
 % xjmp to unseal
-\trgcm{} does not have an instruction for unsealing single sealed capabilities directly, but most sealed capabilities can be achieved unsealed using \texttt{xjmp} and a set of seals with the seal used to seal the sealed capability.
+\trgcm{} does not have an instruction for unsealing single sealed capabilities directly (using the seal that was used to do the sealing), but it can be (partially) simulated using \texttt{xjmp}.
 

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Typo
Typo.

@@
 We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
-This would also allow linear capabilities to have aliases, but only by linear capabilities with disjointed permissions.
+This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.
 
@@
 A component is basically a program with open ends in the form of imports.
-A component also have exports allowing other programs to link with them.
+A component also has exports allowing other programs to link with it.
 We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Typo.

@@
 % Structure
-A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
+A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Typo
Explain the general idea of StkTokens before diving into the details.

@@
 Libraries can be linked together and incomplete programs can be linked with libraries, but two incomplete programs cannot be linked with each other.
-Two components can be linked when their memories, seals, and linear addresses disjointed.
+Two components can be linked when their memories, seals, and linear addresses disjoint.
 They are combined by taking the union of each of their constituents.
@@
 As the examples illustrate, the standard memory and object-like capabilities (sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
-The two properties can, however, be enforced with a calling convention that utilizes the guarantees of linear capabilities.
+The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
+Linear capabilities offer a form of revocation that can be used to prevent this from happening.
 

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
TODO for intro (based on Dominique mail).

--- 2018-07-06 fully-abstract-comp/popl2019/paper.tex
@@
 \label{sec:introduction}
+\todo[inline]{Add a sentence or two about secure compilation to ease the start.}
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
 % KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.
+However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
 
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they have been mentioned in some technical documents.}
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
 % KJAA: Perhaps make it clear by who and when/where?
@@
 \trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main exception.
-In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address space and unbounded integers.
+In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address and unbounded integers.
 \lau{Do we have a good reason to assume an infinite address space for this work?
@@
 % xjmp to unseal
-\trgcm{} does not have an instruction for unsealing single sealed capabilities directly (using the seal that was used to do the sealing), but it can be (partially) simulated using \texttt{xjmp}.
+\trgcm{} does not have an instruction for unsealing single sealed capabilities directly, but most sealed capabilities can be achieved unsealed using \texttt{xjmp} and a set of seals with the seal used to seal the sealed capability.
 
@@
 We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
-This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.
+This would also allow linear capabilities to have aliases, but only by linear capabilities with disjointed permissions.
 
@@
 A component is basically a program with open ends in the form of imports.
-A component also has exports allowing other programs to link with it.
+A component also have exports allowing other programs to link with them.
 We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
typo
Updated step relation to be displayed as inference rule
Added call code and short description.

@@
 \]
-  \caption{The syntax of our capbility machine with seals and linera capbilities.}
+  \caption{The syntax of our capbility machine with seals and linear capbilities.}
   \label{fig:target-syntax}
@@
   \centering
+  \begin{mathpar}
+    \inference{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \\
+        \baddr \le \aaddr \le \eaddr \and \perm \in \{\rwx,\rx\}
+ }{ \Phi \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) }
+              \and
+              \inference{ \neg (\Phi \step \Phi') \and \Phi' \neq \failed}
+                        {\Phi \step \failed }
+  \end{mathpar}
   \[
     \begin{array}{l}
-    \Phi \step
-  \begin{cases}
-    \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) &
-    \arraycolsep=0pt
-      \begin{array}[t]{l}
-        \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \wedge
-        \baddr \le \aaddr \le \eaddr \wedge \perm \in \{\rwx,\rx\}
-      \end{array} \\
-      \failed & \totherwise
-  \end{cases}\\
+  %   \Phi \step
+  % \begin{cases}
+  %   \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) &
+  %   \arraycolsep=0pt
+  %     \begin{array}[t]{l}
+  %       \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \wedge
+  %       \baddr \le \aaddr \le \eaddr \wedge \perm \in \{\rwx,\rx\}
+  %     \end{array} \\
+  %     \failed & \totherwise
+  % \end{cases}\\
   \updPcAddr{\Phi} =
@@
 % Summary of CC
-In the this section, we have described our proposed calling convention, \stktokens{}.
-In this paragraph we summarize it:
-\begin{description}
-\item[Check the base address of the stack capability before and after calls] 
-\item[Make sure that local stack frames are non-empty]
-\item[Create token and data return capability on call] split the stack capability in two to get a stack capability for you local stack frame and a stack capability for the unused part of the stack.
+In the this section, we have described our proposed calling convention, \stktokens{}. The measures taken are:
+\begin{enumerate}
+\item \label{item:check-stkb} Check the base address of the stack capability before and after calls
+\item \label{item:non-empty-sf} Make sure that local stack frames are non-empty
+\item \label{item:return-data} Create token and data return capability on call split the stack capability in two to get a stack capability for you local stack frame and a stack capability for the unused part of the stack.
 The stack capability for your local stack frame is sealed and used for the data part of the return pair.
-\item[Create code return capability on call] Seal the old program pointer.
-\item[Reasonable use of seals] return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
-\end{description}
+\item \label{item:return-code} Create code return capability on call Seal the old program pointer.
+\item Reasonable use of seals return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
+\end{enumerate}
+Item~\ref{item:check-stkb}-\ref{item:return-code} of \stktokens{} are captured by the code in Figure~\ref{fig:call-code} with the exception of checking stack base before calls.
+We do not include this, because it only needs to happen once between two calls which means that the check after a call suffices if the base of the stack pointer is not changed subsequently.
 
@@
   \label{fig:stk-base-abuse}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+{\footnotesize
+\[
+  \begin{array}{r >{$}p{3.5cm}<{$} l >{$}p{3.5cm}<{$} l l}
+    1: & \tmove{\rtmp{1}}{42} & 10: &\tcseal{\rretd}{\rtmp{1}} & 19: &\tcca{\rtmp{2}}{5} \\
+    2: &\tstore{\rstk}{\rtmp{1}} & 11: &\tmove{\rretc}{\pcreg} & 20: &\tjnz{\rtmp{2}}{\rtmp{1}}\\
+    3: &\tcca{\rstk}{(-1)} & 12: & \tcca{\rretc}{5} & 21: &\tcca{\rtmp{2}}{1} \\
+    4: &\tgeta{\rtmp{1}}{\rstk} & 13: & \tcseal{\rretc}{\rtmp{1}} & 22: &   \tjmp{\rtmp{2}} \\
+    5: &\tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}} & 14: & \tmove{\rtmp{1}}{0} & 23: &    \tfail \\
+    6: &\tmove{\rtmp{1}}{\pcreg} & 15: & \txjmp{r_1}{r_2} & 24: &   \tsplice{\rstk}{\rstk}{\rdata} \\
+    7: &\tcca{\rtmp{1}}{(\offpc - 5)} & 16: & \tgetb{\rtmp{1}}{\rstk} & 25: &    \tcca{\rstk}{1}\\
+    8: &\tload{\rtmp{1}}{\rtmp{1}} & 17: &\tminus{\rtmp{1}}{\rtmp{1}}{\stkb} & 26: &    \tmove{\rtmp{2}}{0} \\
+    9: &\tcca{\rtmp{1}}{\offsigma} & 18: &\tmove{\rtmp{2}}{\pcreg} & &
+  \end{array}
+\]
+}
+  \caption{The instructions that constitutes a call.
+The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
+The variable $\offsigma$ is the offset in the set of seals.
+The constant $\stkb$ is the globally agreed on stack base.
+There are a number of ``magic'' numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
+Line 7: $-5$, offset to the first instruction (from the point the capability was moved out of the $\pcreg$-register).
+Line 12: $5$, offset to the return address.
+Line 19: $5$, offset to fail.
+Line 21: offset to address after fail.}
+  \label{fig:call-code}
 \end{figure}

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
More work on section 4. Added more definitions and wrote up about overlay semantics and the oLCM syntax and op sem.

@@
 \]
-  \caption{The syntax of the capability machine with a built-in stack.}
-  \label{fig:src-syntax}
+  \caption{The syntax of \srccm{} with everything new in the overlay semantics in blue. The overlay semantics adds stack pointers, return pointers, and a built-in stack.}
+  \label{fig:source-syntax}
 \end{figure}
 
-The overlay semantics we define for \trgcm{} (\srccm{}) is defined in Figure~\ref{fig:src-syntax} and \ref{fig:src-op-sem}.
-The main things it adds to \trgcm{} is a built-in stack that can be used to store local state as well as a \texttt{call} instruction and a new interpretation of \texttt{xjmp} that corresponds to returning.
 % Describe the things added to the syntax
+The overlay semantics we define for \trgcm{} (\srccm{}) views part of the memory as a built-in stack (Figure~\ref{fig:source-syntax}).
 % Built-in stack
+To this end, it adds a stack and a free stack memory to the executable configurations of \trgcm{}.
+The stack is a list with all the stack frames that are currently inaccessible because they belong to other calls.
+A stack frame contains the stack memory it encapsulates as well as the program point it is supposed to return to.
+The free stack memory is the part of the stack that has not been claimed by a call and thus can be used at this point of time.
 % Stack pointer
+In order to distinguish capabilities for the stack from the capabilities for the rest of the memory, \srccm{} adds stack pointers.
+A stack pointer has a permission, range of authority, and current address, just like capabilities on \trgcm{}, but they are always linear.
 % Return pointers
+The final syntactic constructs added by \srccm{} are the code and data return pointers.
+The data return pointer corresponds to some stack pointer (which in turn corresponds to a linear capability), and the code return pointer corresponds to some capability with read-execute permission.
+Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to.
+They are generated by calls from the capabilities they correspond to, and they are turned back to the capabilities they correspond to upon return (we explain the additions to the operational semantics later).
+Even though they both correspond to capabilities with read permission neither can be used for reading. 
+Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack frame that should be encapsulated at this point which we of course cannot allow.
+
+The opaque nature of the return pointers is reflected in the interpretation of the instructions that are on both \trgcm{} and \srccm{} as none of the capability manipulation instructions on \srccm{} work on them.
+Stack pointers, on the other hand, need to behave just like capabilities, so \srccm{} adds new cases for them in the semantics, e.g.\ \texttt{cca} can now also change the current address of a stack pointer as displayed in Figure~\ref{fig:source-op-sem}.
+Similarly, \texttt{load} and \texttt{store} work on the free part of the stack when working with a stack pointer.
+A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics provide on the underlying machine.
+In other words, it should not be possible to have something else than a stack pointer for that part of memory.
+
+% Back to high-level view
+In \stktokens{} the base address of the memory $\stkb$ used for the stack is agreed upon upfront.
+As \srccm{}'s view adds the stack, it needs to know the base address during execution.
+From the perspective of a well-behaved component, other components can do whatever they want for their calls and local state as long as they cannot break WBCF and LSE for the well-behaved component.
+Due to this, \srccm{} carries a set of trusted addresses $\ta$ which can perform calls in order to push a stack frame to the built-in stack.
+Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
+This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
+The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ($\scall[i]{\offpc,\offsigma}{r_1}{r_2}$ corresponds to the $i$'th instruction in the figure and $\calllen$ is $26$ the number of instructions).
+Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability of course must have the authority to execute the call.
+
+% Call/return
+The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem}, and while it does a fair bit of bookkeeping, it essentially does the following: Given there is a stack pointer in the stack register, $42$ is pushed to the stack to insure that the stack frame will be non-empty.
+Based on the stack pointer, a stack pointer for the callee is created (this corresponds to the return token of \stktokens{}) as well as a return pointer which governs the private part of the stack, i.e.\ addresses $[\aaddr_\stk,\eaddr_\stk]$ which is $42$ and everything (conceptually) below on the stack.
+A new stack frame is pushed to the stack.
+The stack frame consists of local stack (addresses $[\aaddr_\stk,\eaddr_\stk]$) as well as the address where execution is supposed to continue on return.
+The return address in the stack frame is an important piece of meta data as it connects the stack frame to the return point and thus allows us on return to check whether we returned to the right place.
+The return pointers are sealed with the seal dictated by $\offpc$ and $\offsigma$ and placed in registers, so the callee can use them to return.
+Finally, control is transferred in the same fashion \texttt{xjmp} does it.
+
+The return pair can only be used to return with, and as they only live on \srccm in sealed capabilities, it is only natural that \texttt{xjmp} takes care of returning.
+To this end, we add to the semantics of \texttt{xjmp} by redefining $\xjmpres$.
+The function remains the same except when $c_1$ and $c_2$ are code and data return pointers, respectively.
+In this case, the top stack frame is popped from the stack and added to the free stack, and the execution continues at the return point with a stack pointer for the free stack, but only if the code return pointer actually points to the correct return address, the base of the returned stack pointer is $\stkb$, and the returned stack pointer is adjacent to the data return pointer (which corresponds to checking the validity of the return token in \stktokens{}).
 
@@
   \centering
-  \label{fig:src-op-sem}
-  \caption{An excerpt of the operational semantics of the capability machine with a built-in stack.}
+  \begin{mathpar}
+    \inferrule{ \src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
+               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
+               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
+               \src{\perm \in \{\rwx,\rx\}} \\
+               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2}, \dots, \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
+ }
+              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}} }
+  \end{mathpar}
+  \begin{tabular}{|>{$}c<{$}|>{$}p{3.8cm}<{$}|>{\raggedright\arraybackslash}p{6.5cm}|}
+    \hline
+    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
+    \hline
+    \halt                                        & \halted & \\
+    \hline
+    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
+    \hline
+    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{(\perm,\baddr,\eaddr,\aaddr} $ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
+    \hline
+    \cca{r}{\rn}                                 &\srcalt{\updPcAddr{\Phi\updReg{r}{w}}} &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
+    \hline
+    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} & \srcalt{\var{xjmpRes}(c_1,c_2,\Phi')} & $\srcalt{ \Phi' =\var{callConditions}(r_1,r_2,\offpc,\offsigma,\Phi)}$\\
+    \hline 
+    \multicolumn{3}{|c|}{\dots} \\
+    \hline
+    \_                                           & \failed & \totherwise \\
+    \hline
+  \end{tabular}
+
+  \begin{mathpar}
+    \inferrule{
+      \src{r_1 \neq \rtmp{1}}\\
+      \src{\reg(r_1) = \sealed{\sigma,c_1} }\\
+      \src{r_2 \neq \rtmp{1} }\\
+      \src{\reg(r_2) = \sealed{\sigma',c_2} }\\
+      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} }\\
+      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
+      \src{\ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
+      \src{\ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
+      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
+      \src{c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
+      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) }\\
+      \src{\opc = \aaddr + \calllen }\\
+      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }\\
+      \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }\\
+      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} }\\
+      \src{\baddr \leq  \aaddr+\offpc \leq \eaddr }\\
+      \src{\sigma = \sigma_\aaddr + \offsigma }\\
+      \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }\\
+      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)}}\\ 
+      \src{s_c = \sealed{\sigma,c_\opc} }\\
+      \src{s_d = \sealed{\sigma,c_\var{priv\_data}} }\\
+      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}}
+    {\src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}}
+  \end{mathpar}
+\begin{multline*}
+  \srcxjmpres{c_1,c_2,\Phi} = \\
+  \left\{ 
+    \arraycolsep=1.4pt
+    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
+      \arraycolsep=0pt
+    \begin{array}{rl}
+      \Phi & \updReg{\pcreg}{c_1}\\
+           & \updReg{\rdata}{c_2}
+    \end{array}
+    &\phantom{mak} &  $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
+      & & \\
+    \sourcecolor\left.
+      \arraycolsep=0pt
+      \begin{array}[t]{rl}
+        \Phi' & \updReg{\pcreg}{c_\opc}\\
+              & \updReg{\rstk}{c_\stk}\\
+              & \updReg{\rdata}{0}\\
+              & \updReg{\rtmp{1}}{0}\\
+              & \updReg{\rtmp{2}}{0}
+      \end{array}\right.
+    & &
+        \srcalt{$c_1 = \retptrc(\baddr,\eaddr,\opc)$ and
+        $c_2 = \retptrd(\aaddr_\stk,\eaddr_{\stk})$ and
+        $\Phi(\rstk) = \stkptr{\rw,\stkb,\aaddr_\stk-1,\_}$ and
+        $\stkb \leq \aaddr_\stk-1$ and
+        $\Phi = (\mem,\reg,\stkf::\stk,\ms_\stk)$ and
+        $\stkf = (\opc,\ms_{\stk,\priv})$ and
+        $c_\opc = ((\rx,\normal),\baddr,\eaddr,\opc)$ and
+        $\dom(\ms_{\stk,\priv}) = [\aaddr_\stk,\eaddr_{\stk}]$ and
+        $c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk},\aaddr_\stk}$ and
+        $\Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk)$ }\\
+      & & \\
+      \failed &  & \totherwise
+    \end{array} 
+\right.
+\end{multline*}
+  \caption{An excerpt of the operational semantics of \srccm{}.}
+  \label{fig:source-op-sem}
 \end{figure}
 
-% Describe the changes to op sem
-% Step relation change, call case and global constants
-% Call semantics
-% New xjmp semantics (can be defined in terms of a new xjmpresult function!)
-% Example of alternative interpretation for the new syntactic constructs
-
 % Well-formed components
@@
 % Reasonability 
+\todo[inline]{Write about well-formed components and reasonability}
 
 % Full abstraction thm
+\begin{definition}[Source language contextual equivalence]
+  In the source language, we define that $\src{\var{comp}_1 \sconeq \var{comp}_2}$ iff
+  \begin{equation*}
+    \forall \src{\context} \ldotp \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
+  \end{equation*}
+  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$.
+  \dominique{and stkbase?}
+\end{definition}
+
+\begin{definition}[Target language contextual equivalence]
+  In the target language, we define that $\trg{\var{comp}_1 \tconeq \var{comp}_2}$ iff
+  \begin{equation*}
+    \forall \trg{\context} \ldotp \vdash \trg{\context} \Rightarrow \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
+  \end{equation*}
+\end{definition}
+
+\begin{theorem}
+  \label{thm:full-abstraction}
+  For reasonable, well-formed components $\var{comp}_1$ and $\var{comp}_2$, we have
+  \begin{gather*}
+    \src{\var{comp}_1} \sconeq \src{\var{comp}_2}\\
+    \Updownarrow\\
+    \comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}
+  \end{gather*}
+\end{theorem}
 

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
dominique comment improvement.
Theorems and lemmas copied from TR to full-abstractness section.

@@
 \trgcm{} also has an instruction that does the inverse of this, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
-\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities (\texttt{xjmp}).
+\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities (\texttt{xjmp}, see below).
 Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).
@@
 
+\begin{theorem}[FTLR]
+  \label{thm:ftlr}
+  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
+  If
+  \begin{itemize}
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
+  \end{itemize}
+  and one of the following sets of requirements holds:
+  \begin{itemize}
+  \item \begin{itemize}
+    \item $[\baddr,\eaddr] \subseteq \ta$
+    \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
+    \end{itemize}
+  \item
+    \begin{itemize}
+    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
+    \end{itemize}
+  \end{itemize}
+  Then
+  \[
+    \npair{\left(\arraycolsep=0pt\array{l}((\rx,\normal),\baddr,\eaddr,\aaddr),\\
+      ((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lre(W)
+  \]
+\end{theorem}
+
+\begin{lemma}[FTLR for components]
+  \label{lem:ftlr-comps}
+  If
+  \begin{itemize}
+  \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
+  \item $\var{comp}$ is a well-formed component, i.e. $\vdash \var{comp}$
+  \item One of the following holds:
+    \begin{itemize}
+    \item $\dom(\var{comp}.\mscode) \subseteq \ta$ and $\var{comp}$ is a reasonable component (see Section~\ref{sec:reasonability})
+    \item $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
+    \end{itemize}
+  \end{itemize}
+  Then there exists a $W$ such that
+  \begin{itemize}
+  \item $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$
+  \item $\dom(\pwheap)$ can be chosen to not include any finite set of region names 
+  \item $\dom(\pwfree) =
+    \dom(\pwpriv)=\emptyset$
+  \end{itemize}
+\end{lemma}
+
+\begin{lemma}[Adequacy of execution configuration LR]
+  \label{lem:adequacy}
+  If
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$
+  \item $i \leq n$
+  \item $\Phi_S \sterm[i]{\gc}$
+  \end{itemize}
+  then $\Phi_T\term$.
+
+  Also, if
+  \begin{itemize}
+  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$
+  \item $i \leq n$
+  \item $\Phi_T \term[i]$,
+  \end{itemize}
+  then $\Phi_S\sterm{\gc}$.
+\end{lemma}
+
+\begin{lemma}[Compatibility lemma for context plugging]
+  \label{lem:compat-context-plug}
+  If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
+  $\npair[n']{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
+\end{lemma}
+
 \section{Discussion}

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex

@@
 % KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, a potentially expensive operation.
+However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.
 
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they are mentioned here and there by different people.}
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they have been mentioned in some technical documents.}
 % KJAA: Perhaps make it clear by who and when/where?
@@
 A component is basically a program with open ends in the form of imports.
-A component also have exports allowing other programs to link with them.
+A component also has exports allowing other programs to link with it.
 We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.
 % Structure
-A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
+A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.

### 2018-07-06 fully-abstract-comp/popl2019/paper.tex
Added extra explanation of duplication.

@@
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
-Linear capabilities offer a form of revocation that can be used to prevent this from happening.
+Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver no longer has it or a copy of it as it is non-duplicable.} that can be used to prevent this from happening.
 

### 2018-07-07 fully-abstract-comp/popl2019/paper.tex
Possible compilation issue?
Better illustration of sealed capabilities w/ lock symbol in figures.
Intro about well-formedness and reasonability

@@
 \renewcommand{\Cap}{\shareddom{NatTok}}
+\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
+\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
 
 \newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
+\newcommand{\srcxjmpres}[1]{\plainfun{\srcalt{xjmpRes}}{#1}}
 
@@
 \section{Linear Stack and Return Capabilities}
-\todo[inline]{Make sealed capabilities more clear in figures.}
 In this section, we introduce our calling convention \stktokens{} that insures LSE and WBCF.
@@
 
+      \sealsymb{(4.9,3)}
       \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
@@
       % Stack pointer 1
+      \sealsymb{(4.9,5.5)}
       \lincapbrace{(4.6,4)}{(4.6,7)}{adv. stack\\\footnotesize cap. 1}
       % return cap
+      \sealsymb{(4.9,8)}
       \lincapbrace{(4.6,7)}{(4.6,9)}{data return\\\footnotesize cap. 2}
@@
 
+      \sealsymb{(4.9,3)}
       \lincapbrace{(4.6,2)}{(4.6,4)}{}
@@
       \node[opacity=0.5,rotate=-45] at (2.25,10) {\footnotesize Sent to trusted};
+
+      \sealsymb{(4.9,3)}
       \lincapbrace{(4.6,2)}{(4.6,4)}{}
@@
       \node[opacity=0.5,rotate=-45] at (2.25,11) {\footnotesize Returned to adv.};
+      \sealsymb{(4.9,3)}
       \lincapbrace{(4.6,2)}{(4.6,4)}{}
       \lincapbrace{(4.6,4)}{(4.6,7)}{}
+
+      \sealsymb{(4.9,8)}
       \lincapbrace{(4.6,7)}{(4.6,9)}{}
@@
       \lincapbrace{(4.6,2)}{(4.6,7)}{}
+
+      \sealsymb{(4.9,8)}
       \lincapbrace{(4.6,7)}{(4.6,9)}{}
@@
 In this case, the compilation even turns out to be the identity function!
-\todo[inline]{Maybe also introduce reasonability here}
+
+% protection against others purpose of cc
+We can not show that \stktokens{} guarantee much if the security mechanisms of the capability machine are undermined by a poorly configured machine, e.g.\ the same seal is present in different components or the stack capability have aliases.
+It is also possible to for a caller to undermine the security mechanisms during execution.
+For example, the caller could give the return seal for a call to the callee.
+\stktokens{} do not protect a component against itself as the component is assumed to be reasonable. 
+In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that must be satisfied for \stktokens{} to work.
 

### 2018-07-07 fully-abstract-comp/popl2019/paper.tex
Subsections for section 4.
Redefined \comp command, so it can be used for components.

@@
 \renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
+\renewcommand{\comp}{\var{comp}}
 
@@
     \mathrm{export} &\mathrel{::=} s \mapsto w\\
-    \var{comp}_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
-    \var{comp} & \multicolumn{7}{l}{\mathrel{::=} \var{comp}_0 \mid  (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
+    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
+    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
   \end{array}
@@
   \inference{
-    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
-    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
-    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
+    \comp_1 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
+    \comp_2 = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
+    \comp_3 = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
     \mscode[3] = \mscode[1] \uplus \mscode[2] \\
@@
   } {
-    \var{comp}_3 = \var{comp}_1 \bowtie \var{comp}_2
+    \comp_3 = \comp_1 \bowtie \comp_2
   }
@@
   \inference{
-    \var{comp}_0'' = \var{comp}_0 \bowtie \var{comp}_0'
+    \comp_0'' = \comp_0 \bowtie \comp_0'
   }{
-    (\var{comp}_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \var{comp}_0 \bowtie (\var{comp}_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \var{comp}_0'
+    (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \comp_0 \bowtie (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
   }
@@
 % \begin{definition}[Plugging a program into a context]
-%   When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
-%   then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
+%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
+%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
 % \end{definition}
@@
 
+\subsection{Overlay Semantics}
 % Describe the things added to the syntax
@@
 \end{figure}
+\subsection{Well-formedness and reasonability}
 
@@
 
+\subsection{Full abstraction}
 % Full abstraction thm

### 2018-07-07 fully-abstract-comp/popl2019/paper.tex
Added well-defined judgement command local to paper.
More updates related to \comp
Almost finished first draft of section 4. Need to write the ending about full abstraction result.

@@
 \newcommand{\srcxjmpres}[1]{\plainfun{\srcalt{xjmpRes}}{#1}}
+\newcommand{\wdjud}[2][ ]{#1 \vdash #2}
 
@@
 \section{Linear Stack and Return Capabilities}
+\label{sec:stktokens-explained}
 In this section, we introduce our calling convention \stktokens{} that insures LSE and WBCF.
@@
 \subsection{Well-formedness and reasonability}
-
 % Well-formed components
+% \begin{figure}[htb]
+%   \centering
+%   \begin{mathpar}
+%   \inference{
+%     \dom(\mscode) = [\baddr,\eaddr]&
+%     [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
+%     \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
+%     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
+%     \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
+%     A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
+%     \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
+%     \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} & \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}}& \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
+%     \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}\\
+%     \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
+%     \dom(\msdata) \mathrel{\#} \ta
+%   }{
+%     \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+%   }
+%   \and
+%   \inference{
+%     \comp_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
+%     \vdash \comp_0 & (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
+%   }{
+%     \vdash (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
+%   }
+% \end{mathpar}
+%   \caption{Exceprt of well-formedness judgement}
+%   \label{fig:well-formedness}
+% \end{figure}
+% Well-formedness highlights:
+% Main exported
+% Code memory:
+% - Only instructions and seals
+% - calls associated with return seals unique return seal
+% - only uses seals specified for components
+% - (No linear access)
+% Data memory:
+% - Data
+% - Capabilities fora data memory (linearity respected)
+% - Sealed caps sealed with clos seal
+% Exports
+% - Any Word valid in data memory or
+% - Sealed capability for code (sealed with clos seal)
+% Imports
+% - Address must be in data memory
+The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the syntactic requirements necessary to be able to rely on system guarantees.
+For space reasons, we just describe the main points of the judgement and refer to the technical report~\citep{technical_report} for the formal definition.
+For components with a main pair, the main pair must come from the export and the remainder of the component must be well-formed.
+As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
+The return seals $\sigrets$ are the seals supposed to be used for return seals and the closure seals $\sigcloss$ all the other seals.
+If this is not a trusted component, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
+The code memory must only contain set of seals specified by $\sigrets$ and $\sigclos$ and instruction in the form of integers.
+When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), then it must have access to a return seal used by no other call.
+The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
+The capabilities in the data memory must be for the data memory itself, which allow them to form predetermined data structures.
+They can at most have a read-write permission in order to respect Write-XOR-Execute.
+They also need to respect linearity which means that the linear capability cannot have any aliases and it must be for a range of the predetermined linear addresses $A_\linear$.
+The sealed capabilities must use a closure seal and they can seal anything that can reside in data memory.
+The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
+Finally, the addresses of the imports $\overline{\var{import}}$ must be addresses in the data memory.
 
 % Reasonability
-\todo[inline]{Write about well-formed components and reasonability}
+The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially have the measures to use \stktokens{}, but it does not prevent a component from doing something silly that undermines \stktokens{}.
+So in order for \stktokens{} to provide any guarantees for a component, we expect it to not shoot itself in the foot and of course perform the stack check not captured by the call code (Figure~\ref{fig:call-code}).
+We more precisely expect four things of a reasonable component: (1) it checks the stack base address before performing a call.
+As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
+(2) it uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things the do not undermine the security mechanisms \stktokens{} rely on.
+(3) it does not leak return and closure seals or means to get them.
+For instance, this means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module, and it also means that non-sealed capabilities or capabilities sealed with an untrusted seal for code memory cannot be transferred.
+(4) it does not store return and closure seals or means to get them.
+This allow us to safely share capabilities for data memory as they cannot contain seal we do not want to give up.
+The reasonability definition has also been left out for space reasons but can be found in the technical report~\citep{technical_report}.
+
+% \begin{definition}[Reasonable configuration]
+%   \label{def:reasonable-conf}
+%   We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
+%   iff for $n' \leq n$:
+%   \begin{itemize}
+%   \item \emph{Guarantee stack base address before call...}
+%   \item \emph{Use return seals only for calls, use closure seals appropriately...}
+%   \item \emph{Don't store private stuff...} If $\src{\Phi}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
+
+%     Then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ steps.
+%   \item \emph{Don't leak private stuff...}
+%   \end{itemize}
+% \end{definition}
+% Reasonable configuration
+% Reasonable component
 
 \subsection{Full abstraction}
+
+% How a program is lifted to an executable configuration
+Given a program as defined in Definition~\ref{def:program-and-context}, we want to be able to execute it which means that we have to lift it to an executable configuration.
+When we lift a program to an executable configuration, we need to provide it a stack in order for \stktokens{} to work.
+The relation $(\comp_0, c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow \Phi$ lifts a component with a main to an executable configuration.
+It works almost the same on \trgcm{} and \srccm{}.
+On both machines a stack that contains all zeroes is added.
+On \trgcm{} it is added as part of the memory along with the data and code memory.
+On \srccm{} it is added as the free stack in the configuration, and the initial stack is empty as no calls have been made.
+The component needs access to the stack, so a stack pointer is added to the register file in $\rstk$.
+On \trgcm{} this is just a linear read-write capability, but on \srccm{} it is the representation of a stack pointer.
+The entry point of the program is specified by main, so the two capabilities are unsealed (they must have the same seal) and placed in the $\pcreg$ and $\rdata$ registers.
+The remainder of the registers contain zero.
+
+We are now ready to define plugging of components into contexts followed by contextual equivalence:
+%The remaining defintions
+\begin{definition}[Plugging a component into a context]
+  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
+  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
+\end{definition}
+\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
+  \label{def:contextual-equivalence}
+  ~
+  \begin{description}
+  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
+  \begin{equation*}
+    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
+  \end{equation*}
+  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
+  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
+  \begin{equation*}
+    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
+  \end{equation*}
+\end{description}
+\end{definition}
+With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
+We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
+This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
 % Full abstraction thm
-\begin{definition}[Source language contextual equivalence]
-  In the source language, we define that $\src{\var{comp}_1 \sconeq \var{comp}_2}$ iff
-  \begin{equation*}
-    \forall \src{\context} \ldotp \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
-  \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$.
-  \dominique{and stkbase?}
-\end{definition}
-
-\begin{definition}[Target language contextual equivalence]
-  In the target language, we define that $\trg{\var{comp}_1 \tconeq \var{comp}_2}$ iff
-  \begin{equation*}
-    \forall \trg{\context} \ldotp \vdash \trg{\context} \Rightarrow \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
-  \end{equation*}
-\end{definition}
-
 \begin{theorem}
   \label{thm:full-abstraction}
-  For reasonable, well-formed components $\var{comp}_1$ and $\var{comp}_2$, we have
+  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
   \begin{gather*}
-    \src{\var{comp}_1} \sconeq \src{\var{comp}_2}\\
+    \src{\comp_1} \sconeq \src{\comp_2}\\
     \Updownarrow\\
-    \comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}
+    \src{\comp_1} \tconeq \src{\comp_2}
   \end{gather*}
 \end{theorem}
+Readers unfamiliar with full-abstraction may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantee LSE and WBCF.
+\todo[inline]{Write this, i.e.\ explain why full-abstractness prove LSE and WBCF.}
+
+% What does it actually mean?
+% Relies on LSE and WBCF of \srccm - which is obvious from the operational semantics (or defines what it means here?).
 
@@
     % \draw[help lines,yellow] (0,0) grid (10,7);
-    \node at (5,4.7) { ${\src{C_1}\mathrel{\sconeq} \src{C_2}}$ };
-
-    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{C_1}} \sterm[]{\gc}{}}$ };
+    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
+
+    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
     \node at (5,4) { $\mathrel{\Rightarrow}$ };
-    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{C_2}} \sterm[]{\gc}{}}$ };
+    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
 
@@
     \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{C_2}} \cong \comp{{\src{C_2}}}$};
+      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
     \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{C_1}} \cong \comp{{\src{C_1}}}$};
-    % \node at (9,2.7) { $e  {\src{C_1}} \cong \comp{{\src{C_1}}} : tau$ };
-    % \node at (8.7,3.3) { $ {\comp{\trg{\context}}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
-    % \node at (.8,3) { $e  {\src{C_1}} \cong \comp{{\src{C_1}}} : tau$ };
-
-    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{C_1}} \term[]{}}$ };
+      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
+    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
+    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
+    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
+
+    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
     \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
-    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{C_2}} \term[]{}}$ };
-
-    \node at (5,1.3) { ${\src{C_1}}\mathrel{\overset{?}{\tconeq}}{\src{C_2}}$ };
+    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
+
+    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
 
@@
   \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
-  \item $\var{comp}$ is a well-formed component, i.e. $\vdash \var{comp}$
+  \item $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$
   \item One of the following holds:
     \begin{itemize}
-    \item $\dom(\var{comp}.\mscode) \subseteq \ta$ and $\var{comp}$ is a reasonable component (see Section~\ref{sec:reasonability})
-    \item $\dom(\var{comp}.\mscode) \mathrel{\#} \ta$
+    \item $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component (see Section~\ref{sec:reasonability})
+    \item $\dom(\comp.\mscode) \mathrel{\#} \ta$
     \end{itemize}
@@
   \begin{itemize}
-  \item $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$
+  \item $\npair{(\comp,\comp)} \in \lrcomp(W)$
   \item $\dom(\pwheap)$ can be chosen to not include any finite set of region names 

### 2018-07-07 fully-abstract-comp/popl2019/paper.tex
Minor things.

@@
   \label{def:program-and-context}
-  We define a \emph{program} to be a component $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
-
-  A \emph{context} for a component $\var{comp}$ is another component $\var{comp}'$ such that $\var{comp} \bowtie \var{comp}'$ is a program.
+  We define a \emph{program} to be a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
+
+  A \emph{context} for a component $\comp$ is another component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
 \end{definition}
@@
 \end{lemma}
+\subsection{Logical Relation}
+
+\subsection{Component Logical Relation}
+
+\subsection{Context Logical Relation}
 

### 2018-07-07 fully-abstract-comp/technical-report.tex
Comments and corrections to the TR.
Added TODO about well-formedness and trusted addresses in TR.
Redefined \comp command, so it can be used for components.
Minor things.

@@
                                         \withinBounds{\Phi(r_1)} \tand \\
-                                        w_2 = \linCons{\Phi(r_2)} \tand r_2 \neq \pcreg
+                                        w_2 = \linCons{\Phi(r_2)} \tand r_2 \neq \pcreg\\
+                                        \src{\aaddr \in \dom(\Phi.\mem)}
                                       \end{array}
@@
                                         \withinBounds{\Phi(r_1)} \tand \\
-                                        \aaddr \in \dom(\Phi.\ms_\stk) \tand \\ % Lau: this condition seems odd as there should never be a capability for a piece of memory that is unreachable.
+                                        \aaddr \in \dom(\Phi.\ms_\stk) \tand \\
                                         w_2 = \linCons{\Phi(r_2)} \tand r_2 \neq \pcreg
@@
     \begin{array}[t]{l}
-      r_1 \neq \rtmp{1} \tand r_2 \neq \rtmp{1} \\
+      r_1 \neq \rtmp{1} \tand r_2 \neq \rtmp{1} \tand\\
       \Phi(r_1) = \sealed{\sigma_1,c_1} \tand \\
@@
 \end{definition}
+\lau{07-07-2018: TODO, consider if the well-formedness judgement should take the trusted addresses into account, so the trusted component specific things only are required for trusted addresses.
+It already uses $\ta$, so it already ought to be part of the judgement.
+}
+\lau{07-07-2018: TODO: why don't we permit closure seals in data memory? That seems like an oversight.}
 
@@
       \range{\ms_\stk} = \{0\}&
-      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
+      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \uplus \ms_\var{frame} \\
       [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
@@
 \end{definition}
-
-\begin{definition}[Plugging a program into a context]
+\lau{07-07-2018: TODO why don't the above definition case about the stack base?}
+
+\begin{definition}[Plugging a component into a context]
   When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
@@
     \item $\src{\Phi}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$
-
-      \lau{What is this used for?}
     \item $\sigma \in \gsigcloss$ and one of the following holds:
@@
   \begin{equation*}
-    \forall \src{\context} \ldotp \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1},\gsigrets_1,\gsigcloss_1} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2,\gsigrets_2,\gsigcloss_2}}
+    \forall \src{\context} \ldotp \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
   \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and $\gsigcloss_i = \var{comp}_i.\sigcloss$
+  with $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$
+  \lau{06-07-2018: TODO: Dominique, did you resolve this?}
   \dominique{and stkbase?}
@@
     \Updownarrow\\
-    \comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}
+    \src{\var{comp}_1} \tconeq \src{\var{comp}_2}
   \end{gather*}
@@
   \item Consider first the upward arrow.
-    Assume $\comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}$.
+    Assume $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$.
 
@@
 
-    It follows from $\comp{\src{\var{comp}_1}} \tconeq \comp{\src{\var{comp}_2}}$ that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
+    It follows from $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
 

### 2018-07-07 fully-abstract-comp/popl2019/paper.tex
Changes to section 2 after proof reading.

@@
 In this section, we introduce a simple but representative capability machine with linear capabilities.
-We will later introduce a capability machine similar to the one in this section, so to easily distinguish the two, we will call this one \trgcm{} (Linear Capability Machine).
+We will later introduce a capability machine similar to the one in this section, so to easily distinguish the two, we call this one \trgcm{} (Linear Capability Machine).
 \trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main exception.
@@
 % Permissions and Range of authority
-The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}) where $\var{perm}$ is the permission and $[\var{base},\var{end}]$ is the range of authority.
+The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}).
+Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
 \trgcm{} only has memory capabilities, and we limit them to a number of sensible permissions making the available permissions: read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$).
@@
 % Pointer and linearity
-In addition to the two basic things, the capabilities also have a current address $\addr$ and a linearity $\lin$.
+In addition to the two basic things, the capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
 The linearity of a capability can either be $\linear$ or $\normal$ where $\normal$ corresponds to a traditional capability.
 % What is linear capabilities, split/splice
-A linear capability is a capability with no alias that cannot be duplicated.
+A linear capability is a capability that cannot be duplicated.
 This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
 We will see this later in this section when we get to the operational semantics.
+The non-duplicability of linear capabilities means that a linear capability cannot suddenly be aliased if it wasn't to begin with.
 
@@
 % seals
-A set of seals consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$ that is represented by a natural number.
+A set of seals consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$ which is represented by a natural number.
 % sealed caps
 A sealed capability consists of a sealable $\vsc$ and the seal $\sigma$ it has been sealed with.
-The sealables are sets of seals and capabilities, and we refer to the set of them as $\SealableCaps$.
-The system native tokens on \trgcm{} are the sealables and the sealed capabilities and we refer to this set as $\Cap$.
+The sealables are the sets of seals and the capabilities, and we refer to the set of them as $\SealableCaps$.
+The system native tokens $\Cap$ on \trgcm{} are the sealables and the sealed capabilities.
 
@@
 The words on \trgcm{} are the system native tokens in $\Cap$ and data which is represented by $\ints$.
-We assume a finite set $\RegName$ of register names that at least contains registers $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as a function from register names to words.
-A memory is a full memory and a memory segment is part of a memory.
-Both are defined as functions from addresses, defined as the natural numbers, to words, but the memory segment functions only need to be partial.
-$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution with an attempt to break the capability machine guarantees.
+We assume a finite set of register names $\RegName$ that at least contains register names $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
+We have a notion of a complete memory where all memory addresses contains a word and a memory segment that need not contain words on all addresses.
+Both are represented as functions from addresses, defined as the natural numbers, to words, but the memory segment functions only need to be partial.
+$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.\ an attempt to break the capability machine guarantees.
 An executable configuration is a register file and memory pair.
@@
 \]
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.4cm}<{$}|>{\raggedright\arraybackslash}p{7cm}|}
+  \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.8cm}|}
     \hline
@@
     \hline
-    \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{\rn,r}{w_1,w_2}} & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\rn}$ \\
-    \hline
-    \load{r_1}{r_2}                              & \updPcAddr{\Phi\updReg{r_1}{w_1}\update{\mem.\aaddr}{w_\aaddr}} & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $w_a = \linCons{w_1}$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$\\
-    \hline
-    \store{r_1}{r_2}                             & \updPcAddr{\Phi\updReg{r_2}{w_2}\update{\mem.\aaddr}{\Phi(r_2)}} & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
+    \move{r}{\rn}                                & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\rn}$ \\
+    \hline
+    \load{r_1}{r_2}                              & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{w_1}\\ &\update{\mem.\aaddr}{w_\aaddr})\endarray & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$ and $w_a = \linCons{w_1}$\\
+    \hline
+    \store{r_1}{r_2}                             & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ & \update{\mem.\aaddr}{\Phi(r_2)})\endarray & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
     \hline
@@
     \hline
-    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{\Phi\updReg{r_3,r_1,r_2}{w,c_1,c_2}} & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_1)}$\\
-    \hline
-    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{\Phi\updReg{r_2,r_3,r_1}{w_2,w_3,c}}& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_1)}$\\
+    \hline
+    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
     \hline
@@
   \end{tabular}
-\caption{An excerpt of the operational semantics of our capability machine with sealing and linear capabilities.}
+\caption{An excerpt of the operational semantics of \trgcm{}}
   \label{fig:target-op-sem}
@@
 The instructions that can be considered standard are: unconditional and a conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability machine specific instructions simply inspect system native tokens or project one of their fields.
+The simplest of the capability machine specific instructions simply inspect the type of system native tokens or project one of their fields.
 These are \texttt{gettype} (for getting the type of a system token) and \texttt{getp}, \texttt{getl}, \texttt{getb}, \texttt{gete}, and \texttt{geta} (for projecting permission, linearity, base, end, and current address(/seal), respectively).
-There are two instructions for manipulating the current address(/seal) of a capability(/set of seals): \texttt{cca} changes the current address(/seal) of a capability(/set of seals) by some offset, and \texttt{seta2b} set the current address(/seal) of a capability(/set of seals) to the base address(/seal).
+There are two instructions for manipulating the current address(/seal) of a capability(/set of seals): \texttt{cca} changes the current address(/seal) of a capability(/set of seals) by some offset, and \texttt{seta2b} sets the current address(/seal) of a capability(/set of seals) to the base address(/seal).
 The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
-Generally speaking, a capability machine needs an instruction for reducing the range of authority of an capability.
-Because \trgcm{} has linear capabilities the instruction for this splits the capability in two rather than reducing the range of authority (instruction \texttt{split}).
-\trgcm{} also has an instruction that does the inverse of this, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
-\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities (\texttt{xjmp}, see below).
+Generally speaking, a capability machine needs an instruction for reducing the range of authority of a capability.
+Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two rather than reducing the range of authority (\texttt{split}).
+\trgcm{} also has an instruction that does the inverse of splitting, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
+\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities by crossing security boundaries (\texttt{xjmp}, see below).
 Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).
@@
 The executed instruction is determined by the capability in $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
-In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the range of addresses where the capability can operate.
+In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
 If both things are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relatively to $\Phi$.
-The interpretation of some of the instructions are also displayed in Figure~\ref{fig:target-op-sem}.
+The interpretation of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
 In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
@@
 It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
-This means that when a linear capability is moved, the source register should be cleared.
-To take care of this and prevent duplicating the operational semantics, we use the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
-When the \texttt{move} instruction is performed, the source register $r_2$ is overwritten with $\linCons{\Phi(r_2)}$ which clears the register if the contents was linear and leaves it unchanged otherwise.
+When a linear capability is moved, the source register must be cleared, so the capability is not duplicated.
+This is done uniformly in the semantics using two approaches.
+In the first one, we use the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
+When a word $w$ that can possibly be a linear capability is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
+In the case of \texttt{move}, the source register $r_2$ is overwritten with $\linCons{\Phi(r_2)}$.
+We see the other approach later.
 
@@
 The \texttt{store} and \texttt{load} instructions are fairly standard: They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority, and they clear the source of a linear capability transfer.
-Linear capabilities introduce one complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
+Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
 Clearing a memory address corresponds to writing $0$ to the memory, but in order to write to memory we need to provide a capability with write-permission.
@@
 The \texttt{cca} instruction changes the current address or seal of a capability or set of seals, respectively, by a given offset.
-By making an in-place update, the instruction does not need to take special care in order to maintain linearity.
+It handles linearity by updating the capability in-place, i.e.\ the source register is used as the target register.
+This is the second approach we use to uniformly handle linearity.
 The \texttt{jmp} instruction is a fairly simple jump that just sets the $\pcreg$-register.
@@
 Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
+This also makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
+The opaque sealed capabilities encapsulate a security domain's local state and authority, and it only becomes accessible again when control is transferred to the security domain.
 Some care should be taken when sealing as reuse of the same seal for multiple closures would make it possible to jump to the code of one capability with the environment of another.
@@
 Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
-For normal capabilities, reduction of authority can be done without actually giving up any authority if the capability is duplicated first.
+For normal capabilities, reduction of authority can be done without actually giving up any authority by duplicating the capability first.
 With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
 In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
-The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as it range of authority.
+The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as its range of authority.
 Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
-With \texttt{split} we can reduce the range of authority of a linear capability without losing any authority as we retain it in the other capability.
-However, if we only have \texttt{split}, then in some sense splitting a capability is not completely lossless as we may have authority over the same range of addresses, but we have that authority in two capabilities rather than just one.
+With \texttt{split}, we can reduce the range of authority of a linear capability without losing any authority as we retain it in the second capability.
+However, if we only have \texttt{split}, then in some sense splitting a capability is not completely lossless as we may have authority over the same range of addresses after a \texttt{split}, but we have that authority in two capabilities rather than just one.
 In order to make up for this, we have a \texttt{splice} instruction that essentially does the inverse of \texttt{split}.
@@
 At this time, the executable configuration is our only way to specify the state of the capability machine.
-We would, however, much prefer a notion of a program that only describe the essential parts of an entity on the machine and only later instrument the program with whatever necessary to make it into an executable configuration.
+We would, however, much prefer a notion of programs that only describes the essential parts of an entity on the machine and only later instrument the program to make it into an executable configuration.
 In later sections, we describe different entities on the capability machine that interact with each other.
-In order for them to interact, we need to be able to link them together and to this end, we introduce a notion of a component.
-A component is basically a program with open ends in the form of imports.
-A component also has exports allowing other programs to link with it.
+In order for them to interact, we need to be able to link them together, and to this end we introduce a notion of components.
+A component is basically a program with open ends in the form of imports that need to be satisfied.
+A component also has exports allowing other components to link with it.
 We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.
 % Structure
-A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in later sections.}, and a set of all the linear addresses.
+A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses.
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
-The exports are words associated with a symbol.
+The exports are words each associated with a symbol.
 %base = library
 A base component can be seen as library implementations because it does not specify a main entry point of execution.
-Components are either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
+A component is either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
 % Linking .
@@
 The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
-In Figure~\ref{fig:target-component-and-linking}, $\bowtie$ defines linking.
+Components and linking are defined in Figure~\ref{fig:target-component-and-linking}.
 \todo[inline]{Add better separation between component and linking in figure.}
@@
 % write xor exec
-In Section~\ref{sec:form-secur-with}, we will get back to what a program is instrumented with in order to create an executable configuration. \todo[inline]{Make sure we got back to this!}
-We will, however, at this point mention that all programs will be instrumented with the Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+In Section~\ref{sec:form-secur-with}, we will get back to how a program is instrumented to create an executable configuration.
+We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
@@
 \]
+\vspace{0.75cm}
 \begin{mathpar}

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
First draft of full-abstraction subsection.

@@
 \subsection{Full abstraction}
+All that is left before we state the full-abstraction theorem is to define how programs are lifted to executable configurations and contextual equivalence for each of the two capability machines we have defined.
 
@@
 
-We are now ready to define plugging of components into contexts followed by contextual equivalence:
+Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
 %The remaining defintions
@@
 \end{description}
+where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
 \end{definition}
@@
 \end{theorem}
-Readers unfamiliar with full-abstraction may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantee LSE and WBCF.
-\todo[inline]{Write this, i.e.\ explain why full-abstractness prove LSE and WBCF.}
-
-% What does it actually mean?
-% Relies on LSE and WBCF of \srccm - which is obvious from the operational semantics (or defines what it means here?).
+Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
+Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
+This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
+If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
+In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
+In order to enforce these properties, components on \trgcm{} use \stktokens{}.
+Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
 

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
Pass over section 4 (except full-abstraction sub section)

@@
 \end{enumerate}
-Item~\ref{item:check-stkb}-\ref{item:return-code} of \stktokens{} are captured by the code in Figure~\ref{fig:call-code} with the exception of checking stack base before calls.
-We do not include this, because it only needs to happen once between two calls which means that the check after a call suffices if the base of the stack pointer is not changed subsequently.
+Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} with the exception of checking stack base before calls.
+We do not include this check because it only needs to happen once between two calls which means that the check after a call suffices if the base of the stack pointer is not changed subsequently.
 
@@
 A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
-Our proof uses a novel approach called full abstract overlay semantics.
+Our proof uses a novel approach we call fully abstract overlay semantics.
 In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
 An overlay semantics is basically a different view on an existing semantics.
-For instance, it may have new syntactic constructs and have special semantics for these things, but they will correspond to something on the underlying machine.
-In the overlay semantics we are about to define for \trgcm{}, there is an explicit stack pointer construct on the machine, but on the underlying semantics the stack pointer corresponds directly to a linear capability for the part of memory the overlay semantics has deemed to be the stack.
+For instance, it may have new syntactic constructs and have special semantics for these things, but they correspond to something on the underlying machine.
+In the overlay semantics we are about to define for \trgcm{}, an explicit stack pointer construct is part of the new view.
+The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
 When we have defined the overlay semantics, we prove full-abstractness for a compilation from the overlay semantics to the underlying semantics.
-Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as it is really just a different view on the same thing. 
-In this case, the compilation even turns out to be the identity function!
+Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.
+In our case, the compilation even turns out to be the identity function!
 
@@
 For example, the caller could give the return seal for a call to the callee.
-\stktokens{} do not protect a component against itself as the component is assumed to be reasonable. 
-In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that must be satisfied for \stktokens{} to work.
-
-\begin{figure}[htb]
+\stktokens{} do not protect a component against itself as the component is assumed to behave reasonably.
+In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components from unsafe behavior.
+
+\begin{figure}[b]
   \centering
@@
     \begin{array}{rcl}
-      \src{\SealableCaps} & \defbnf& \SealableCaps \mid \src{\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}} \mid \src{\retptrd(\basebnf,\aendbnf)}\\
-                          & &  \mid \src{\retptrc(\basebnf,\aendbnf,\addrbnf)}\\
+      \src{\SealableCaps} & \defbnf& \SealableCaps \mid \src{\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}} \mid\\
+                          & &   \src{\retptrd(\basebnf,\aendbnf)} \mid \src{\retptrc(\basebnf,\aendbnf,\addrbnf)}\\
       \multicolumn{3}{c}{
@@
 \]
-  \caption{The syntax of \srccm{} with everything new in the overlay semantics in blue. The overlay semantics adds stack pointers, return pointers, and a built-in stack.}
+\caption{The syntax of \srccm{}.
+  \srccm{} extends \trgcm{} by adding stack pointers, return pointers, and a built-in stack.
+  Everything specific to the overlay semantics is written in blue.
+}
   \label{fig:source-syntax}
@@
 % Built-in stack
-To this end, it adds a stack and a free stack memory to the executable configurations of \trgcm{}.
-The stack is a list with all the stack frames that are currently inaccessible because they belong to other calls.
-A stack frame contains the stack memory it encapsulates as well as the program point it is supposed to return to.
+To this end, it adds a call stack and a free stack memory to the executable configurations of \trgcm{}.
+The call stack is a list with all the stack frames that are currently inaccessible because they belong to other calls.
+A stack frame contains encapsulated stack memory as well as the program point execution is supposed to return to.
 The free stack memory is the part of the stack that has not been claimed by a call and thus can be used at this point of time.
@@
 The data return pointer corresponds to some stack pointer (which in turn corresponds to a linear capability), and the code return pointer corresponds to some capability with read-execute permission.
-Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to.
-They are generated by calls from the capabilities they correspond to, and they are turned back to the capabilities they correspond to upon return (we explain the additions to the operational semantics later).
-Even though they both correspond to capabilities with read permission neither can be used for reading. 
-Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack frame that should be encapsulated at this point which we of course cannot allow.
-
-The opaque nature of the return pointers is reflected in the interpretation of the instructions that are on both \trgcm{} and \srccm{} as none of the capability manipulation instructions on \srccm{} work on them.
+Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to on the underlying machine.
+Return pointers are generated by calls from the capabilities they correspond to, and they are turned back to the capabilities they correspond to upon return.
+Even though the return pointers both correspond to capabilities with read permission, neither can be used for reading. 
+Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack memory that is encapsulated in a stack frame which would break LSE.
+
+The opaque nature of the return pointers is reflected in the interpretation of the instructions common to both \trgcm{} and \srccm{} as \srccm{} does not add special interpretation for them.
 Stack pointers, on the other hand, need to behave just like capabilities, so \srccm{} adds new cases for them in the semantics, e.g.\ \texttt{cca} can now also change the current address of a stack pointer as displayed in Figure~\ref{fig:source-op-sem}.
-Similarly, \texttt{load} and \texttt{store} work on the free part of the stack when working with a stack pointer.
-A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics provide on the underlying machine.
+Similarly, \texttt{load} and \texttt{store} work on the free part of the stack when provided with a stack pointer.
+A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics has on the underlying machine.
 In other words, it should not be possible to have something else than a stack pointer for that part of memory.
@@
 In \stktokens{} the base address of the memory $\stkb$ used for the stack is agreed upon upfront.
-As \srccm{}'s view adds the stack, it needs to know the base address during execution.
+As \srccm{}'s view builds-in the stack, it needs to know the base address during execution.
 From the perspective of a well-behaved component, other components can do whatever they want for their calls and local state as long as they cannot break WBCF and LSE for the well-behaved component.
-Due to this, \srccm{} carries a set of trusted addresses $\ta$ which can perform calls in order to push a stack frame to the built-in stack.
+Due to this, \srccm{} has a set of trusted addresses $\ta$ that are allowed to perform machine calls that push stack frames to the call stack.
+To make these constants available during execution, they are passed around in the \trgcm{} step relation.
 Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
 This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
-The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ($\scall[i]{\offpc,\offsigma}{r_1}{r_2}$ corresponds to the $i$'th instruction in the figure and $\calllen$ is $26$ the number of instructions).
-Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability of course must have the authority to execute the call.
+The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ({\footnotesize $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is $26$, i.e.\ the number of instructions).
+Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.
 
 % Call/return
-The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem}, and while it does a fair bit of bookkeeping, it essentially does the following: Given there is a stack pointer in the stack register, $42$ is pushed to the stack to insure that the stack frame will be non-empty.
-Based on the stack pointer, a stack pointer for the callee is created (this corresponds to the return token of \stktokens{}) as well as a return pointer which governs the private part of the stack, i.e.\ addresses $[\aaddr_\stk,\eaddr_\stk]$ which is $42$ and everything (conceptually) below on the stack.
+The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem}, and it may look daunting due to a fair bit of bookkeeping, but it essentially does the following: Given a stack pointer in the stack register, $42$ is pushed to the stack to insure that the stack frame will be non-empty.
+Based on the stack pointer in $\rstk$, a stack pointer for the callee is created (this corresponds to the return token of \stktokens{}) as well as a return pointer which governs the private part of the stack, i.e.\ addresses $[\aaddr_\stk,\eaddr_\stk]$ which is $42$ and everything (conceptually) below on the stack.
 A new stack frame is pushed to the stack.
-The stack frame consists of local stack (addresses $[\aaddr_\stk,\eaddr_\stk]$) as well as the address where execution is supposed to continue on return.
+The stack frame consists of the local stack (addresses $[\baddr_\stk,\aaddr_\stk-1]$) as well as the address where execution is supposed to continue on return.
 The return address in the stack frame is an important piece of meta data as it connects the stack frame to the return point and thus allows us on return to check whether we returned to the right place.
-The return pointers are sealed with the seal dictated by $\offpc$ and $\offsigma$ and placed in registers, so the callee can use them to return.
+The return pointers are sealed with the seal dictated by $\offpc$ and $\offsigma$ and placed in registers $\rretc$ and $\rretd$, so the callee can use them to return.
 Finally, control is transferred in the same fashion \texttt{xjmp} does it.
 
-The return pair can only be used to return with, and as they only live on \srccm in sealed capabilities, it is only natural that \texttt{xjmp} takes care of returning.
-To this end, we add to the semantics of \texttt{xjmp} by redefining $\xjmpres$.
+The return pair can only be used to return with.
+They only exists on \srccm{} in sealed capabilities, so it is only natural that \texttt{xjmp} takes care of returning.
+To this end, we update the semantics of \texttt{xjmp} by redefining $\xjmpres{}$.
 The function remains the same except when $c_1$ and $c_2$ are code and data return pointers, respectively.
-In this case, the top stack frame is popped from the stack and added to the free stack, and the execution continues at the return point with a stack pointer for the free stack, but only if the code return pointer actually points to the correct return address, the base of the returned stack pointer is $\stkb$, and the returned stack pointer is adjacent to the data return pointer (which corresponds to checking the validity of the return token in \stktokens{}).
+In this case, the top stack frame is popped from the stack and added to the free stack.
+The execution continues at the return point with a stack pointer in $\rstk$ for the free stack that now includes the memory of the popped stack frame.
+However, the execution only continues if the code return pointer actually points to the correct return address, the base of the returned stack pointer is $\stkb$, and the returned stack pointer is adjacent to the data return pointer (which corresponds to checking the validity of the return token in \stktokens{}).
+
+By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack which means that it insures LSE.
+Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack which means that we have WBCF.
+In other words by inspection of the operational semantics, \srccm{} has LSE and WBCF.
+It is important to point this out as \stktokens{} guarantee exactly the LSE and WBCF we have on \srccm{}.
 
@@
 For space reasons, we just describe the main points of the judgement and refer to the technical report~\citep{technical_report} for the formal definition.
-For components with a main pair, the main pair must come from the export and the remainder of the component must be well-formed.
+For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
 As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
-The return seals $\sigrets$ are the seals supposed to be used for return seals and the closure seals $\sigcloss$ all the other seals.
-If this is not a trusted component, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
-The code memory must only contain set of seals specified by $\sigrets$ and $\sigclos$ and instruction in the form of integers.
-When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), then it must have access to a return seal used by no other call.
+The return seals $\sigrets$ are the seals supposed to be used to seal return pointers and the closure seals $\sigcloss$ all the other seals in a component .
+If a component is not trusted, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
+The code memory may contain sets of seals, but only the seals in $\sigrets$ and $\sigclos$.
+Other than that, code memory only contains instructions in the for of integers.
+When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), it must have access to the return seal specified by that call.
+The return seal also needs to be unique to that call, so no other call can specify the same seal.
 The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
-The capabilities in the data memory must be for the data memory itself, which allow them to form predetermined data structures.
-They can at most have a read-write permission in order to respect Write-XOR-Execute.
-They also need to respect linearity which means that the linear capability cannot have any aliases and it must be for a range of the predetermined linear addresses $A_\linear$.
-The sealed capabilities must use a closure seal and they can seal anything that can reside in data memory.
+The capabilities in the data memory can only have authority over the data memory itself.
+By allowing this, a component can be constructed with predetermined data structures.
+In order to respect Write-XOR-Execute, the capabilities can at most have a read-write permission.
+The capabilities also need to respect linearity which means that a linear capability cannot have any aliases, and it must be for a range of the predetermined linear addresses $A_\linear$.
+The sealed capabilities must be sealed with a closure seal, and the sealable can be anything that is allowed to reside in data memory.
 The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
@@
 % Reasonability
-The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially have the measures to use \stktokens{}, but it does not prevent a component from doing something silly that undermines \stktokens{}.
-So in order for \stktokens{} to provide any guarantees for a component, we expect it to not shoot itself in the foot and of course perform the stack check not captured by the call code (Figure~\ref{fig:call-code}).
-We more precisely expect four things of a reasonable component: (1) it checks the stack base address before performing a call.
+The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
+In order for \stktokens{} to provide any guarantees for a component, we expect it to not shoot itself in the foot and of course perform the necessary stack check not captured by the call code (Figure~\ref{fig:call-code}).
+We more precisely expect four things of a reasonable component: (1) It checks the stack base address before performing a call.
 As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
-(2) it uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things the do not undermine the security mechanisms \stktokens{} rely on.
-(3) it does not leak return and closure seals or means to get them.
-For instance, this means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module, and it also means that non-sealed capabilities or capabilities sealed with an untrusted seal for code memory cannot be transferred.
-(4) it does not store return and closure seals or means to get them.
-This allow us to safely share capabilities for data memory as they cannot contain seal we do not want to give up.
-The reasonability definition has also been left out for space reasons but can be found in the technical report~\citep{technical_report}.
+(2) It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
+(3) It does not leak return and closure seals or means to retrieve them.
+This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
+There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
+(4) It does not store return and closure seals or means to get them.
+By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
+The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.
 

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
Write a discussion about practical implementatibility of lincaps.

--- 2018-07-07 fully-abstract-comp/popl2019/paper.tex
@@
 \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
-\item Performance?  Implementability of linear caps?
 \item Sharing stack references accross component boundaries is supported
@@
 \end{itemize}
+
+\subsection{Practical applicability}
+We believe there are good arguments for practical applicability of \stktokens{}.
+The strong security guarantees are proven in a way that is reusable as part of a bigger proof of compiler security.
+Its costs are
+\begin{itemize}
+\item a constant and limited amount of checks on every boundary crossing.
+\item possibly a small memory overhead because every stack frame must be of non-zero length
+\end{itemize}
+The main caveat is that we rely on the assumption that capability machines like CHERI can be extended with linear capabilities in an efficient way.
+
+Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
+As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like $\tmove{}$, $\tstore{}$ and $\tload{}$.
+Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
+Our instructions are a bit clumsy from this point-of-view because, for example, $\tmove{}$ or $\tstore{}$ will zero the source register resp. memory location if the value being written is linear.
+A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
+Adding splice and split will also consume some opcode space.
+
+Another problem is caused by the move semantics for $\tload{}$ in the presence of multiple hardware threads.
+In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
+This means that a $\tload{}$ of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
+This is in principle not a problem except that a compare-and-swap is significantly slower than a regular $\tload{}$ (on the order of 10x slower or more).
+When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
+Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
+If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than $\tload$'s worst-case...
 

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
A note

--- 2018-07-08 fully-abstract-comp/popl2019/paper.tex
@@
 \begin{itemize}
-\item mention that tail calls are supported through xjmp
+\item mention that tail calls are supported through xjmp (perhaps mention earlier: when introducing scall?)
 \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
-\item Performance?  Implementability of linear caps?
 \item Sharing stack references accross component boundaries is supported

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
Another note about performance of loads.
Subsection formatting.
Comment.

@@
 
-\subsection{Overlay Semantics}
+\subsection{Overlay semantics}
 % Describe the things added to the syntax
@@
 Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
+\todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}
 
@@
 If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than $\tload$'s worst-case...
+The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
 

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
Added "wish list" of definitions for section 5.

@@
 \label{sec:fa-proof}
-
-\begin{itemize}
-\item Logical relation
-\item FTLR
-\item Sketch high-level structure of the proof
-\end{itemize}
-
+\todo[inline]{Write small introduction.
+Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
+% \begin{itemize}
+% \item Logical relation
+% \item FTLR
+% \item Sketch high-level structure of the proof
+% \end{itemize}
+\todo[inline]{Describe figure here or save it for proof sketch subsection?}
 \begin{figure}
@@
 
-\begin{theorem}[FTLR]
+\subsection{Logical relation}
+\label{subsec:worlds}
+The corner stone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
+\todo[inline]{write a bit more introduction}
+
+Due to space considerations, we cannot give a comprehensive walk through of the logical relation used for the full-abstraction proof.
+We refer to~\citet{skorstengaard_reasoning_2017} for a more comprehensive walk through of a logical relation for a capability machine and refer to the technical report~\citep{technical_report} for the details of the logical relation used here.
+The main differences between the logical relation defined here and the one by~\citet{skorstengaard_reasoning_2017} are: this is a binary, cross-language logical relation, the worlds are significantly different
+\begin{theorem}
+  \label{thm:recursive-domain-eq}
+  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ and there exists an isomorphism $\xi$ such that
+  \[
+    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
+  \]
+  and for $\hat{W},\hat{W}' \in \Wor$
+  \[
+    \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})
+  \]
+\end{theorem}
+
+Where $\Worlds$, $\Worldh$, and $\Worldfs$ are defined as follows
+\[
+  \Worldh = \RegionName \parfun (\Regions + \Regionh)
+\]
+and
+\[
+  \Worlds = \RegionName \parfun (\Regions \times \Addr)
+\]
+and
+\[
+  \Worldfs = \RegionName \parfun \Regions
+\]
+where $\RegionName = \nats$.
+\begin{multline*}
+  \Regionh = 
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
+\end{multline*}
+and
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
+    \{\revoked\}
+  \end{array} \right.
+\]
+% Describe purepart \oplus and \uplus
+
+\begin{multline*}
+  \lrexj(W) =
+  \\\left\{\begin{array}{l}
+      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
+      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}
+  \end{multline*}
+
+  \begin{multline*}
+    \lre(W)= \\
+    \left\{ \begin{array}{l}
+      \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} | \\
+      \quad\forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}} \tand\\
+      \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}} \tand W \oplus W_\lrrs \oplus W_\lrm\\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}
+  \end{multline*}
+\[
+  \lrol =\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+  \forall i \leq n \ldotp
+    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}
+\]
+\[
+  \lror = \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+  \forall i \leq n \ldotp 
+    \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
+\]
+
+\begin{multline*}
+  \lrv[\gc=(\ta,\stkb,\gsigrets,\gsigcloss)](W) = \\
+  \begin{array}[t]{l}
+%   \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+%
+    \left\{
+    \begin{array}{l}
+      \npair{\left(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T} \right)} \; | \; \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
+      \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
+      \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\ \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
+      % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
+      % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+    \end{array}
+    \right\}\cup\\
+    \left\{
+    \begin{array}{l}
+    \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)}\; | \; [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
+      \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+    \end{array}
+    \right\} \cup \\
+    \left\{
+    \begin{array}{l}
+      \npair{\left(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr) \right)} \; | \;
+        \perm \not\in \{\rx,\rwx\} \tand \\
+        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in
+        % \lrp(W) \wedge \\
+        \quad\perm \in \{\ro,\rw\} \Rightarrow \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \tand \\
+        \quad\perm = \rw  \Rightarrow \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
+    \end{array}
+    \right\} \cup \\ \dots
+%
+    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    % \begin{array}{l}
+    %   [b,e] \mathrel{\#} \ta \tand\\
+    %   \begin{array}{r l l }
+    %     % \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in
+    %     % \lrp(W) \wedge \\
+    %     \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
+    %     \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
+    %     % we are excluding rwx pointers.
+    %     \perm \neq \rwx \wedge \\
+    %     % \perm = \rwx &\Rightarrow&
+    %     % \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
+    %     % \wedge \\
+    %     % \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
+    %     \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
+    %     \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
+    %     \lin = \normal \\ \endarray
+    %   \end{array}
+    % \end{array}
+    % \right\}
+  \end{array}
+\end{multline*}
+
+\begin{multline*}
+  \lrvtrusted[\square,\gc=(\ta,\stkb,\gsigrets,\gsigcloss)](W) = \lrv(W)\cup \\
+  \begin{array}[t]{l}
+    \left\{ \begin{array}{l} \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} 
+    \; | \; \exists r \in \dom(\pwheap) \ldotp \\
+              \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc} \tand \dom(\code) \subseteq \ta \\
+              \quad \tand [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \tand \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss
+    \end{array}
+    \right\} \cup \\
+    \left\{
+    \begin{array}{l}
+      \npair{\left(\src{((\perm,\normal),\baddr,\eaddr,\aaddr)},((\perm,\normal),\baddr,\eaddr,\aaddr) \right)} \; |  \\
+      \quad \perm \sqsubseteq \rx \tand 
+       [\baddr,\eaddr] \subseteq \ta \tand 
+       \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
+    \end{array}
+    \right\}
+  \end{array}
+\end{multline*}
+
+\begin{multline*}
+  \stackWriteCond{W} =\\
+  \left\{
+    \begin{array}{l}\npair{A} \mid 
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
+      \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
+      \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\end{multline*}
+\[
+  \xReadCond{W} = \left\{ \npair{A} \middle|
+    \begin{array}{l}
+      \exists r \in \addressable{\normal, \pwheap} \ldotp \\
+       \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\tand
+       \dom(\code) \supseteq A 
+    \end{array}
+  \right\}
+\]
+
+\[
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
+H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc,
+H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+\]
+
+\todo[inline]{$H^\mathrm{code}$ not that interesting, cut for space.}
+\begin{multline*}
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
+  \left\{
+    \begin{array}{l}
+      \npair{(\code \uplus \mspad, \code \uplus \mspad)} \mid
+      \dom(\code) = [\baddr,\eaddr] \tand  \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \tand\\
+      ([\baddr - 1, \eaddr + 1] \subseteq \ta \tand
+      \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \tor \\
+      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \tand \\
+%      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
+      \forall a \in \dom(\code)\ldotp \npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
+    \end{array}
+  \right\}
+\end{multline*}
+
+\begin{multline*}
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  \begin{array}[t]{l}
+\left\{
+    \begin{array}{l}
+      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr'+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \mid 
+        \sigrets \subseteq \gsigrets \tand \\
+        \quad\dom(\code) \subseteq \ta \tand 
+        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \quad\aaddr = \aaddr' + \retoffset \tand \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
+        \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
+        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
+    \end{array}
+      \right\} \cup \\
+\left\{
+    \begin{array}{l}
+      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid
+        \sigrets \subseteq \gsigrets \tand 
+        \dom(\code) \subseteq \ta \tand\\
+        \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
+        \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
+        \qquad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \qquad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
+    \end{array}
+    \right\} \\
+  \end{array}\\
+    \text{ if } \sigma \in \sigrets\\
+\end{multline*}
+and
+\begin{multline*}
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  \left\{
+    \begin{array}{l}
+ \npair{(\vsc, \vsc' )} \mid
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
+        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
+         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
+         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
+    \end{array}
+  \right\}\\
+  \text{ if } \sigma \in \sigcloss
+\end{multline*}
+
+\begin{lemma}[FTLR]
   \label{thm:ftlr}
@@
   \]
-\end{theorem}
+\end{lemma}
+
+\subsection{Component logical relation}
+
+\begin{multline*}
+  \lrcomp(W) =\\
+  \left\{\begin{aligned}
+      &\npair{\var{comp},\var{comp}} \;\mid \;\\
+      &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+      &\qquad\text{For all } W' \future W \ldotp \\
+      &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+      &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
+      &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+    \end{aligned}
+  \right\}\\
+\cup \left\{
+    \begin{multlined}
+      \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
+      \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
+      \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
+    \end{multlined}
+  \right\} 
+\end{multline*}
 
@@
 \end{lemma}
-
-\begin{lemma}[Adequacy of execution configuration LR]
-  \label{lem:adequacy}
-  If
-  \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$
-  \item $i \leq n$
-  \item $\Phi_S \sterm[i]{\gc}$
-  \end{itemize}
-  then $\Phi_T\term$.
-
-  Also, if
-  \begin{itemize}
-  \item $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$
-  \item $i \leq n$
-  \item $\Phi_T \term[i]$,
-  \end{itemize}
-  then $\Phi_S\sterm{\gc}$.
-\end{lemma}
-
 \begin{lemma}[Compatibility lemma for context plugging]
@@
 \end{lemma}
-\subsection{Logical Relation}
-
-\subsection{Component Logical Relation}
-
-\subsection{Context Logical Relation}
+
+\subsection{Executable configuration logical relation}
+
+\begin{multline*}
+  \lrec[\square,\gc = (\ta,\stkb)](W) = \\
+\left\{
+  \begin{array}{l}
+     \npair{\left(
+    (\ms_S,\reg_S,\stk,\ms_\stk),
+    (\ms_T,\reg_T)\right)} \mid \\
+    \quad \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
+    \qquad \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
+    \qquad \reg_S(\pcreg) \neq \retptrc(\_) \tand
+    \reg_S(\rdata) \neq \retptrd(\_) \tand \\
+    \qquad \nonExec{\reg_S(\rdata)} \tand  
+     \nonExec{\reg_T(\rdata)} \tand \\
+    \qquad \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
+  \end{array}
+\right\}
+\end{multline*}
+\begin{lemma}[Adequacy of execution configuration LR]
+  \label{lem:adequacy}
+  If $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$ and
+ $i \leq n$ and
+and $\Phi_S \sterm[i]{\gc}$,
+  then $\Phi_T\term$.
+
+  Also, if
+ $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$ and
+ $i \leq n$
+and $\Phi_T \term[i]$,
+  then $\Phi_S\sterm{\gc}$.
+\end{lemma}
+
+\subsection{Proof sketch}
+\begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]~
+  % \item Consider first the upward arrow.
+  %   Assume $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$.
+
+  %   Take a $\src{\context}$ such that $\vdash \src{\context}$, take $\src{\ta[,i]}
+  %   = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and
+  %   $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$ and we will prove that
+  %   $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}} \Leftrightarrow
+  %   \src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
+
+  %   By symmetry, we can assume w.l.o.g. that $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$ and prove that $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
+  %   Note that this implies that $\src{\context}$ is a valid context for both $\src{\var{comp}_1}$ and $\src{\var{comp}_2}$.
+
+  %   First, we show that also $\plug{\context}{\var{comp}_1} \trg{\term}$.
+  %   Take $n$ the amount of steps in the termination of $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$.
+  %   It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\preceq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+  %   It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+  %   Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\preceq,\gc_1](W_1\uplus W_1')$
+  %   Together with $\src{\plug{\context}{\var{comp}_1} \sterm[n]{\gc_1}}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \trg{\term}$.
+
+  %   It follows from $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
+
+  %   It now remains to show that also $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
+  %   Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \trg{\term}$.
+  %   It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\succeq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+  %   It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+  %   Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\succeq,\gc_2](W_2\uplus W_2')$
+  %   Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$, concluding this direction of the proof.
+
+  First consider the right arrow:
+
+    Assume $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$.
+Take $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$.
+%
+    Take a $\trg{\context}$ such that $\vdash \trg{\context}$ and we will prove that
+    $\trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow
+    \trg{\plug{\context}{\var{comp}_2} \term}$.
+%
+    By symmetry, we can assume w.l.o.g. that $\trg{\plug{\context}{\var{comp}_1} \term}$ and prove that $\trg{\plug{\context}{\var{comp}_2} \term}$.
+    Note that this implies that $\trg{\context}$ is a valid context for both $\trg{\var{comp}_1}$ and $\trg{\var{comp}_2}$.
+%
+    First, we show that also $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
+    Take $n$ the amount of steps in the termination of $\plug{\context}{\var{comp}_1} \trg{\term}$.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
+    Together with $\plug{\context}{\var{comp}_1} \trg{\term[n]}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
+%
+    It follows from $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
+%
+    It now remains to show that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
+    Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
+    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$
+    Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_2} \trg{\term}$, concluding the second direction of the proof.
+
+ The left arrow is proven in a similar manner.
+  \end{itemize}
+\end{proof}
 

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
commands and formatting.
Formatting.
Restructuring of section 5's LR subsection
Text about worlds and subsubtitles for the subsection

@@
 \renewcommand{\comp}{\var{comp}}
+\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
+\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
 
@@
   \begin{gather*}
-    \src{\comp_1} \sconeq \src{\comp_2}\\
-    \Updownarrow\\
-    \src{\comp_1} \tconeq \src{\comp_2}
+    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2}
   \end{gather*}
@@
 The main differences between the logical relation defined here and the one by~\citet{skorstengaard_reasoning_2017} are: this is a binary, cross-language logical relation, the worlds are significantly different
-\begin{theorem}
+
+For the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but not so interesting.
+
+\subsubsection{Worlds}
+Generally speaking, worlds are used in a logical relation to model the memory of a language.
+\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds should model this by being triples of sub-worlds.
+All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to some regions that depend on the kind of sub-world.
+where the regions differ for each of the sub-worlds.
+Generally, the regions consist of a world indexed relation that specifies the memories allowed by the region and a region type.
+The region type indicates how the region can be addressed which is important for linearity.
+We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
+The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
+The $\spatial$ region may seem odd as it not addressable, but all the relations in the logical relation are world indexed.
+If the register-file relation and the memory relation both have worlds with the same ownership, then they can both contain linear capabilities that address the same memory.
+If a spatial region is owned by the register-file relation, then it still has to be modelled in the memory which it can use a spatial region for.
+To capture this, we define a relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
+We also define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
+In a sense, linear capabilities are revocable, so the sub-worlds that allow spatial regions must also have a $\revoked$ region to model revocation.
+Finally, the $\pure$ region also contains a seal interpretation function.
+The seal interpretation function is world indexed, and it specifies what sealables a given seal can be used to seal.
+We split the regions in two categories ($\Wor$ given below):
+\begin{multline*}
+  \Regionh = 
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
+\end{multline*}
+\begin{multline*}
+  \Regions = 
+    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup \\
+    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
+    \{\revoked\}
+\end{multline*}
+and define the sub-worlds as follows:
+\begin{align*}
+  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
+  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
+  \Worldfs &={} \RegionName \parfun \Regions
+\end{align*}
+The world for the memory can contain every kind of region because it is addressable by both normal and linear capabilities.
+The call stack and free stack should only be addressable by linear capabilities, so they only allow regions from $\Regions$.
+Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.
+
+The worlds $\World$ are defined as $(\Worldh \times \Worlds \times \Worldfs)$ and they are recursive as they contain world indexed relations\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
+Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct them in the lemma below.
+\begin{lemma}
   \label{thm:recursive-domain-eq}
@@
   \]
-  and for $\hat{W},\hat{W}' \in \Wor$
+  and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
+\end{lemma}
+The future world relation $\future$ is fairly standard as it is extensional and requires regions to stay the same with the exception that $\spatial$ regions can become $\spatialo$ or $\revoked$ in a future world.
+% \begin{multline*}
+%   \lrexj(W) =
+%   \\\left\{\begin{array}{l}
+%       \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
+%       \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
+%       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+%       \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+%       \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+%     \end{array}
+%     \right\}
+% \end{multline*}
+
+\subsubsection{Code region}
 \[
-    \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
+H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc,
+H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
 \]
-\end{theorem}
-
-Where $\Worlds$, $\Worldh$, and $\Worldfs$ are defined as follows
-\[
-  \Worldh = \RegionName \parfun (\Regions + \Regionh)
-\]
-and
-\[
-  \Worlds = \RegionName \parfun (\Regions \times \Addr)
-\]
-and
-\[
-  \Worldfs = \RegionName \parfun \Regions
-\]
-where $\RegionName = \nats$.
+
+\todo[inline]{Describe $H^\mathrm{code}$}
+% \begin{multline*}
+%   H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
+%   \left\{
+%     \begin{array}{l}
+%       \npair{(\code \uplus \mspad, \code \uplus \mspad)} \mid
+%       \dom(\code) = [\baddr,\eaddr] \tand  \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \tand\\
+%       ([\baddr - 1, \eaddr + 1] \subseteq \ta \tand
+%       \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \tor \\
+%       \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \tand \\
+% %      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
+%       \forall a \in \dom(\code)\ldotp \npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
+%     \end{array}
+%   \right\}
+% \end{multline*}
+
 \begin{multline*}
-  \Regionh = 
-  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
-  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  \begin{array}[t]{l}
+\left\{
+    \begin{array}{l}
+      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr'+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \mid 
+        \sigrets \subseteq \gsigrets \tand \\
+        \quad\dom(\code) \subseteq \ta \tand 
+        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \quad\aaddr = \aaddr' + \retoffset \tand \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
+        \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
+        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
+    \end{array}
+      \right\} \cup \\
+\left\{
+    \begin{array}{l}
+      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid
+        \sigrets \subseteq \gsigrets \tand 
+        \dom(\code) \subseteq \ta \tand\\
+        \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
+        \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
+        \qquad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \qquad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
+    \end{array}
+    \right\} \\
+  \end{array}\\
+    \text{ if } \sigma \in \sigrets\\
 \end{multline*}
 and
-\[
-  \Regions = \left\{
+\begin{multline*}
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  \left\{
     \begin{array}{l}
-    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
-    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
-    \{\revoked\}
-  \end{array} \right.
-\]
-% Describe purepart \oplus and \uplus
-
-\begin{multline*}
-  \lrexj(W) =
-  \\\left\{\begin{array}{l}
-      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
-      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
-      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-    \end{array}
-    \right\}
+ \npair{(\vsc, \vsc' )} \mid
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
+        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
+         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
+         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
+    \end{array}
+  \right\}\\
+  \text{ if } \sigma \in \sigcloss
 \end{multline*}
 
+\subsubsection{Expression relations}
   \begin{multline*}
@@
 
+\subsubsection{Value relations}
 \begin{multline*}
@@
 \]
-
-\[
-  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
-H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc,
-H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
-\]
-
-\todo[inline]{$H^\mathrm{code}$ not that interesting, cut for space.}
-\begin{multline*}
-  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
-  \left\{
-    \begin{array}{l}
-      \npair{(\code \uplus \mspad, \code \uplus \mspad)} \mid
-      \dom(\code) = [\baddr,\eaddr] \tand  \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \tand\\
-      ([\baddr - 1, \eaddr + 1] \subseteq \ta \tand
-      \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \tor \\
-      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \tand \\
-%      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
-      \forall a \in \dom(\code)\ldotp \npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
-    \end{array}
-  \right\}
-\end{multline*}
-
-\begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
-  \begin{array}[t]{l}
-\left\{
-    \begin{array}{l}
-      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr'+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \mid 
-        \sigrets \subseteq \gsigrets \tand \\
-        \quad\dom(\code) \subseteq \ta \tand 
-        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \quad\aaddr = \aaddr' + \retoffset \tand \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
-        \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
-        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
-    \end{array}
-      \right\} \cup \\
-\left\{
-    \begin{array}{l}
-      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid
-        \sigrets \subseteq \gsigrets \tand 
-        \dom(\code) \subseteq \ta \tand\\
-        \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
-        \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
-        \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
-        \qquad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \qquad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
-    \end{array}
-    \right\} \\
-  \end{array}\\
-    \text{ if } \sigma \in \sigrets\\
-\end{multline*}
-and
-\begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
-  (\ta,\stkb) \; \sigma \; \hat{W}= \\
-  \left\{
-    \begin{array}{l}
- \npair{(\vsc, \vsc' )} \mid
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
-         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
-         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
-    \end{array}
-  \right\}\\
-  \text{ if } \sigma \in \sigcloss
-\end{multline*}
 
@@
 \subsection{Proof sketch}
-\begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]~
+\begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
+  \todo[inline]{Rewrite}
   % \item Consider first the upward arrow.
@@
  The left arrow is proven in a similar manner.
-  \end{itemize}
 \end{proof}

### 2018-07-08 fully-abstract-comp/popl2019/paper.tex
More work on section 5.

@@
 \label{subsec:worlds}
-The corner stone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
-\todo[inline]{write a bit more introduction}
-
-Due to space considerations, we cannot give a comprehensive walk through of the logical relation used for the full-abstraction proof.
-We refer to~\citet{skorstengaard_reasoning_2017} for a more comprehensive walk through of a logical relation for a capability machine and refer to the technical report~\citep{technical_report} for the details of the logical relation used here.
-The main differences between the logical relation defined here and the one by~\citet{skorstengaard_reasoning_2017} are: this is a binary, cross-language logical relation, the worlds are significantly different
-
-For the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but not so interesting.
+The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
+\todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
+Unfortunately due to space considerations, we cannot give a comprehensive description of it.
+Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensible description of a logical relation for a capability machine.
+Finally, for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.
 
@@
 Finally, the $\pure$ region also contains a seal interpretation function.
-The seal interpretation function is world indexed, and it specifies what sealables a given seal can be used to seal.
+The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
 We split the regions in two categories ($\Wor$ given below):
@@
 The future world relation $\future$ is fairly standard as it is extensional and requires regions to stay the same with the exception that $\spatial$ regions can become $\spatialo$ or $\revoked$ in a future world.
+
+\subsubsection{Code region}
+The code region is a central region to the logical relation used to model the code memory of a component.
+It is especially interesting because it specifies what the return and closure seals in a component can be used to seal.
+The code region is defined as follows:
+\[
+  \codereg{\sigrets,\sigcloss,\mscode,\gc} \defeq (\pure,
+  H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \mscode \; \gc,
+  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)
+\]
+It is a $\pure$ region which makes it persistent in future worlds.
+The memory relation $H^\mathrm{code,\square} \; (\dots)$ is very selective and only accepts the pair $(\mscode,\mscode)$ and only on the following conditions: The code memory $\mscode$ must be well-formed according to $\vdash$.
+Further, all the words contained in $\mscode$ must be non-linear and in the value relation, but as we will see later we have two value relations, a trusted and an untrusted, and the value relation used depends on whether $\mscode$ is trusted.
+For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have.
+The trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.
+If the addresses of $\mscode$ are in $\ta$, then the trusted value relation is used, and if $\mscode$ is disjoint with $\ta$, then the untrusted value relation is used.
+Further, if $\ta$ is untrusted, then $\sigrets$ has to be empty which means that $\mscode$ cannot contain return seals.
+
+% \todo[inline]{Describe $H^\mathrm{code}$}
 % \begin{multline*}
-%   \lrexj(W) =
-%   \\\left\{\begin{array}{l}
-%       \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
-%       \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
-%       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-%       \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
-%       \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-%     \end{array}
-%     \right\}
-% \end{multline*}
-
-\subsubsection{Code region}
-\[
-  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
-H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc,
-H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
-\]
-
-\todo[inline]{Describe $H^\mathrm{code}$}
-% \begin{multline*}
-%   H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
+%   H^\mathrm{code} \; \sigrets \; \sigcloss \; \mscode \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
 %   \left\{
 %     \begin{array}{l}
-%       \npair{(\code \uplus \mspad, \code \uplus \mspad)} \mid
-%       \dom(\code) = [\baddr,\eaddr] \tand  \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \tand\\
+%       \npair{(\mscode \uplus \mspad, \mscode \uplus \mspad)} \mid
+%       \dom(\mscode) = [\baddr,\eaddr] \tand  \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \tand\\
 %       ([\baddr - 1, \eaddr + 1] \subseteq \ta \tand
@@
 % %      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
-%       \forall a \in \dom(\code)\ldotp \npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
+%       \forall a \in \dom(\mscode)\ldotp \npair{(\mscode(a),\mscode(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
 %     \end{array}
@@
 
+\todo[inline]{Describe $H^\mathrm{code,\square}_\sigma$}
+
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
   \begin{array}[t]{l}
@@
         \sigrets \subseteq \gsigrets \tand \\
-        \quad\dom(\code) \subseteq \ta \tand 
-        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \quad\aaddr = \aaddr' + \retoffset \tand \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
+        \quad\dom(\mscode) \subseteq \ta \tand 
+        \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
         \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
@@
         \sigrets \subseteq \gsigrets \tand 
-        \dom(\code) \subseteq \ta \tand\\
+        \dom(\mscode) \subseteq \ta \tand\\
         \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
@@
         \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
-        \qquad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \qquad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
+        \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
     \end{array}
@@
 \end{multline*}
-and
-\begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
-  (\ta,\stkb) \; \sigma \; \hat{W}= \\
-  \left\{
-    \begin{array}{l}
- \npair{(\vsc, \vsc' )} \mid
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
-        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
-         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
-         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
-    \end{array}
-  \right\}\\
-  \text{ if } \sigma \in \sigcloss
-\end{multline*}
+We leave out the formal definition of $H^\mathrm{code,\square}_\sigma \; (\dots)$ for $\sigma \in \sigcloss$ and provide an informal description.
+What closure seals can be used for depends on whether the code is trusted.
+If the code is untrusted, i.e.\ $\dom(\mscode) \# \ta$, then the seal can be used for everything in the untrusted value relation can be sealed which correspond to everything the code has access to.
+On the other hand, if the code is trusted, then it depends on whether the sealable in question is executable.
+If it is executable, then it can be in the trusted value relation because we want to allow sealing of capabilities for the trusted code; but if it isn't executable, then the sealable has to be in the untrusted value relation.
+
+% and
+% \begin{multline*}
+%   H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \;
+%   (\ta,\stkb) \; \sigma \; \hat{W}= \\
+%   \left\{
+%     \begin{array}{l}
+%  \npair{(\vsc, \vsc' )} \mid
+%         (\dom(\mscode) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
+%         (\dom(\mscode) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
+%          \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
+%          \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
+%     \end{array}
+%   \right\}\\
+%   \text{ if } \sigma \in \sigcloss
+% \end{multline*}
 
@@
 \begin{multline*}
-    \lre(W)= \\
-    \left\{ \begin{array}{l}
-      \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} | \\
-      \quad\forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}} \tand\\
-      \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}} \tand W \oplus W_\lrrs \oplus W_\lrm\\
-      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+  \lrexj(W) =
+  \\\left\{\begin{array}{l}
+      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
+      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
     \end{array}
@@
 \end{multline*}
+
+  % \begin{multline*}
+  %   \lre(W)= \\
+  %   \left\{ \begin{array}{l}
+  %     \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} | \\
+  %     \quad\forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+  %     \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
+  %     \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}} \tand\\
+  %     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}} \tand W \oplus W_\lrrs \oplus W_\lrm\\
+  %     \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+  %   \end{array}
+  %   \right\}
+  % \end{multline*}
 \[
@@
 \subsubsection{Value relations}
+  \todo[inline]{Do we cut the sealed case of the untrusted value relation?}
+
 \begin{multline*}

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Added paragraph about tail calls to section 4.

@@
 However, the execution only continues if the code return pointer actually points to the correct return address, the base of the returned stack pointer is $\stkb$, and the returned stack pointer is adjacent to the data return pointer (which corresponds to checking the validity of the return token in \stktokens{}).
+
+\srccm{} supports tail calls.
+A tail call is a call from a caller that is done executing and thus don't need to be returned to.
+This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in call.
+To perform a tail call, the caller simply transfers control to the callee using \texttt{xjmp}.
+The callee needs to return to the caller's caller, so the caller leaves the return pair they received for the callee to use.
 
@@
 \begin{itemize}
-\item mention that tail calls are supported through xjmp (perhaps mention earlier: when introducing scall?)
 \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Wrote the rest of code region subsubsection.

@@
 \]
-It is a $\pure$ region which makes it persistent in future worlds.
+The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
 The memory relation $H^\mathrm{code,\square} \; (\dots)$ is very selective and only accepts the pair $(\mscode,\mscode)$ and only on the following conditions: The code memory $\mscode$ must be well-formed according to $\vdash$.
@@
 
+the seal interpretation function $h^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
+the return seals can only be used to seal return pointers, but only when they actually represent meaning full returns.
+for instance, a code return pointer should actually be associated with a trusted call, i.e.\ it should point to the address just after a call.
+the corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to do a couple of things to complete the call safely.
+further, the call they are associated with has to be the call that uses the seal in question.
+the code return pointer is not linear which means that it can be reused which is fine as it can only be used to return from the same call.
+on the other hand, the data return pointers are linear which is what prevents reuse of a return pair.
+multiple nested calls from the same trusted call will result in multiple stack frames that are each sealed with the same return seal.
+the seal interpretation function should allow a data return pointer for any of those pointers, so it simply requires that there exists a region $\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}$ (a region that only accepts $(\ms_S,\ms_T)$ and requires the contents to be in the untrusted value relation) in the part of the world that governs the call stack.
+The data return pointer should point to that stack frame.
+A region in the world for the call stack is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
+\todo[inline]{Consider writing something about how this relates to prove? We have to show that there is a stack frame, but we can rely on it when we receive a capability with the correct seal.}
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb,\gsigrets,\gsigcloss) \; \sigma \; \hat{W}= \\
   \begin{array}[t]{l}
@@
   \end{array}\\
-    \text{ if } \sigma \in \sigrets\\
+    \text{ if } \sigma \in \sigrets
 \end{multline*}

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Comments and a few notes for contributions.
Finished expression relation subsubsection.

@@
 \renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
+\renewcommand{\Worlds}{\World_{\mathrm{call\ stack}}}
 
@@
 \newcommand{\srccm}{\textsc{oLCM}}
+\newcommand{\extend}[1]{}
 
@@
 
-\todo[inline]{Add itemized contributions}
-
-\paragraph{Outline} Blabla
+\todo[inline]{Add itemized contributions (see comments). Outline?}
+% The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure local state encapsulation and well-bracketed control flow, and (2) a novel proof technique that we call fully-abstract overlay semantics.
+
+% Presents the first formal model of a capability machine with linear capabilities (Section 2)
+% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
+% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
+
+% \begin{itemize}
+% \item 
+% \end{itemize}
+
+%\paragraph{Outline} Blabla
 
@@
 % Full abstraction thm
+\todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
 \begin{theorem}
@@
 
-\subsubsection{Expression relations}
-\begin{multline*}
-  \lrexj(W) =
-  \\\left\{\begin{array}{l}
-      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
-      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
-      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-    \end{array}
-    \right\}
-\end{multline*}
-
+\subsubsection{Expression relations and friends}
+The expression relation $\mathcal{E}$ in this logical relation contains pairs of capabilities that can safely be used for execution with respect to the give world.
+That is when capabilities are plugged into configurations with related register files  and related memory relative to the world, then the executions observably behave the same way.
+% the other expression relation
+With sealed capabilities in the mix, we will also need to be able to say something similar to the above but with a specific data capability in the register file.
+To this end, we have the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$ that relates two pairs of capabilities and requires that they are safe to use for the \texttt{xjmp} instruction.
   % \begin{multline*}
@@
   % \end{multline*}
-\[
-  \lrol =\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+\begin{multline*}
+  \lrexj(W) =
+  \\\left\{\begin{array}{l}
+      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
+      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
+      \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}
+\end{multline*}
+% register file relation
+The register-files relation $\lrr(\{\rdata\}) (W_\lrrs )$ relates register files when their registers, with register $\rdata$ and $\pcreg$ as the exception, contain words related in the untrusted value relation.
+% memory relation
+A memory on \trgcm{} $\ms_T$ is related to a heap, call stack and free stack on \srccm{} if the following is possible: $\ms_T$ can be split in three disjoint parts.
+One part relates to $\ms_S$ with respect to the heap world in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
+The second part should relate to the free stack $\ms_\stk$ also in the standard way, but with respect to the free stack world.
+The final part relates to the call stack $\stk$ which means that it can be further split into parts that each are related to one of the stack frames of $\stk$ with respect to the call stack world.
+The partition must be done in a way that preserves the order of the stack frames in $\stk$.
+% , so the top stack frame relates to a part of memory where all the other parts are above in the address space, the second stack frame relates to a part of memory where all the other parts are above in the address space with the exception of the topmost part, and so on.
+% Further, all stack frames should be in addresses above $\stkb$.
+\extend{Write about the particulars of the three memory satisfaction: heap relation makes sure only one interpretation for each seal. Free stack $\stkb$}
+The \texttt{xjmp} expression relation requires register files and memories related under different worlds, namely $W_\lrrs$ and $W_\lrm$, that are related with $\oplus$.
+This means that the two worlds have the same regions, but distribute the ownership of partial regions between them.
+By doing so, they make sure that for any part of memory only one of them can have a linear capability it.
+The two pairs of capabilities related by $\lrexj$ are plugged into their respective configurations using the $\xjmpres{}$ function, so it corresponds to how an $\texttt{xjmp}$ would insert them.
+% observation relation
+Finally, the observably accepted behaviour is defined by the observation relation $\lro$.
+\begin{align*}
+  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
   \forall i \leq n \ldotp
-    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}
-\]
-\[
-  \lror = \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
+  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
   \forall i \leq n \ldotp 
     \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
-\]
+\end{align*}
+The only observable behavior of the capability machines is termination. 
+This means that the observation relations say that if one configuration terminates successfully, then the other one must terminate successfully as well.
+The fact that the observation relation is split into two parts means that we really have a logical approximation where one machine approximates the behavior of the other machine.
 

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Formatting of FTLR, so it takes up less space.

@@
   For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
-  If
-  \begin{itemize}
-  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
-  \end{itemize}
-  and one of the following sets of requirements holds:
-  \begin{itemize}
-  \item \begin{itemize}
-    \item $[\baddr,\eaddr] \subseteq \ta$
-    \item $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps.
-    \end{itemize}
-  \item
-    \begin{itemize}
-    \item $[\baddr,\eaddr] \mathrel{\#} \ta$
-    \end{itemize}
-  \end{itemize}
-  Then
+  If $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
+  and either $[\baddr,\eaddr] \subseteq \ta$ and $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
+$[\baddr,\eaddr] \mathrel{\#} \ta$,
+then 
   \[
-    \npair{\left(\arraycolsep=0pt\array{l}((\rx,\normal),\baddr,\eaddr,\aaddr),\\
-      ((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \in \lre(W)
+    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr),
+      ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W)
   \]

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
typo

@@
     \hline
-    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{(\perm,\baddr,\eaddr,\aaddr} $ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
+    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
     \hline

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
minor comments and eddits.

--- 2018-07-09 fully-abstract-comp/popl2019/paper.tex
@@
 \renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
-\renewcommand{\Worlds}{\World_{\mathrm{call\ stack}}}
 
@@
 \newcommand{\srccm}{\textsc{oLCM}}
-\newcommand{\extend}[1]{}
 
@@
 
-\todo[inline]{Add itemized contributions (see comments). Outline?}
-% The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure local state encapsulation and well-bracketed control flow, and (2) a novel proof technique that we call fully-abstract overlay semantics.
-
-% Presents the first formal model of a capability machine with linear capabilities (Section 2)
-% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
-% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
-
-% \begin{itemize}
-% \item 
-% \end{itemize}
-
-%\paragraph{Outline} Blabla
+\todo[inline]{Add itemized contributions}
+
+\paragraph{Outline} Blabla
 
@@
 %%% Seals
-Any reasonable capability machine needs a way to setup boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
+% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
+
+Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
@@
 This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
-However, \stktokens{} is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
+However, \stktokens{} is a modular calling scheme: no restriction is put on adversarial components that we do not expect trusted components to meet.
 This means that any component can ensure WBCF and LSE by employing \stktokens{}.
@@
 \item \label{item:non-empty-sf} Make sure that local stack frames are non-empty
-\item \label{item:return-data} Create token and data return capability on call split the stack capability in two to get a stack capability for you local stack frame and a stack capability for the unused part of the stack.
+% KJAA: There's a comma (or something!) missing below. 
+\item \label{item:return-data} Create token and data return capability on call, split the stack capability in two to get a stack capability for your local stack frame, and a stack capability for the unused part of the stack.
 The stack capability for your local stack frame is sealed and used for the data part of the return pair.
+  % KJAA: Same with the last two items
 \item \label{item:return-code} Create code return capability on call Seal the old program pointer.
@@
 The code memory may contain sets of seals, but only the seals in $\sigrets$ and $\sigclos$.
-Other than that, code memory only contains instructions in the for of integers.
+Other than that, code memory only contains instructions in the form of integers.
 When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), it must have access to the return seal specified by that call.
@@
 % How a program is lifted to an executable configuration
-Given a program as defined in Definition~\ref{def:program-and-context}, we want to be able to execute it which means that we have to lift it to an executable configuration.
+Given a program as defined in Definition~\ref{def:program-and-context}, we want to be able to execute it, meaning we have to lift it to an executable configuration.
 When we lift a program to an executable configuration, we need to provide it a stack in order for \stktokens{} to work.
@@
 % Full abstraction thm
-\todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
 \begin{theorem}
@@
 \]
-The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
+It is a $\pure$ region which makes it persistent in future worlds.
 The memory relation $H^\mathrm{code,\square} \; (\dots)$ is very selective and only accepts the pair $(\mscode,\mscode)$ and only on the following conditions: The code memory $\mscode$ must be well-formed according to $\vdash$.
@@
 
-the seal interpretation function $h^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
-the return seals can only be used to seal return pointers, but only when they actually represent meaning full returns.
-for instance, a code return pointer should actually be associated with a trusted call, i.e.\ it should point to the address just after a call.
-the corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to do a couple of things to complete the call safely.
-further, the call they are associated with has to be the call that uses the seal in question.
-the code return pointer is not linear which means that it can be reused which is fine as it can only be used to return from the same call.
-on the other hand, the data return pointers are linear which is what prevents reuse of a return pair.
-multiple nested calls from the same trusted call will result in multiple stack frames that are each sealed with the same return seal.
-the seal interpretation function should allow a data return pointer for any of those pointers, so it simply requires that there exists a region $\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}$ (a region that only accepts $(\ms_S,\ms_T)$ and requires the contents to be in the untrusted value relation) in the part of the world that governs the call stack.
-The data return pointer should point to that stack frame.
-A region in the world for the call stack is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
-\todo[inline]{Consider writing something about how this relates to prove? We have to show that there is a stack frame, but we can rely on it when we receive a capability with the correct seal.}
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb,\gsigrets,\gsigcloss) \; \sigma \; \hat{W}= \\
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb) \; \sigma \; \hat{W}= \\
   \begin{array}[t]{l}
@@
   \end{array}\\
-    \text{ if } \sigma \in \sigrets
+    \text{ if } \sigma \in \sigrets\\
 \end{multline*}
@@
 
-\subsubsection{Expression relations and friends}
-The expression relation $\mathcal{E}$ in this logical relation contains pairs of capabilities that can safely be used for execution with respect to the give world.
-That is when capabilities are plugged into configurations with related register files  and related memory relative to the world, then the executions observably behave the same way.
-% the other expression relation
-With sealed capabilities in the mix, we will also need to be able to say something similar to the above but with a specific data capability in the register file.
-To this end, we have the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$ that relates two pairs of capabilities and requires that they are safe to use for the \texttt{xjmp} instruction.
+\subsubsection{Expression relations}
+\begin{multline*}
+  \lrexj(W) =
+  \\\left\{\begin{array}{l}
+      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
+      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}
+\end{multline*}
+
   % \begin{multline*}
@@
   % \end{multline*}
-\begin{multline*}
-  \lrexj(W) =
-  \\\left\{\begin{array}{l}
-      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
-      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
-      \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-    \end{array}
-    \right\}
-\end{multline*}
-% register file relation
-The register-files relation $\lrr(\{\rdata\}) (W_\lrrs )$ relates register files when their registers, with register $\rdata$ and $\pcreg$ as the exception, contain words related in the untrusted value relation.
-% memory relation
-A memory on \trgcm{} $\ms_T$ is related to a heap, call stack and free stack on \srccm{} if the following is possible: $\ms_T$ can be split in three disjoint parts.
-One part relates to $\ms_S$ with respect to the heap world in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
-The second part should relate to the free stack $\ms_\stk$ also in the standard way, but with respect to the free stack world.
-The final part relates to the call stack $\stk$ which means that it can be further split into parts that each are related to one of the stack frames of $\stk$ with respect to the call stack world.
-The partition must be done in a way that preserves the order of the stack frames in $\stk$.
-% , so the top stack frame relates to a part of memory where all the other parts are above in the address space, the second stack frame relates to a part of memory where all the other parts are above in the address space with the exception of the topmost part, and so on.
-% Further, all stack frames should be in addresses above $\stkb$.
-\extend{Write about the particulars of the three memory satisfaction: heap relation makes sure only one interpretation for each seal. Free stack $\stkb$}
-The \texttt{xjmp} expression relation requires register files and memories related under different worlds, namely $W_\lrrs$ and $W_\lrm$, that are related with $\oplus$.
-This means that the two worlds have the same regions, but distribute the ownership of partial regions between them.
-By doing so, they make sure that for any part of memory only one of them can have a linear capability it.
-The two pairs of capabilities related by $\lrexj$ are plugged into their respective configurations using the $\xjmpres{}$ function, so it corresponds to how an $\texttt{xjmp}$ would insert them.
-% observation relation
-Finally, the observably accepted behaviour is defined by the observation relation $\lro$.
-\begin{align*}
-  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+\[
+  \lrol =\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
   \forall i \leq n \ldotp
-    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
-  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}
+\]
+\[
+  \lror = \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
   \forall i \leq n \ldotp 
     \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
-\end{align*}
-The only observable behavior of the capability machines is termination. 
-This means that the observation relations say that if one configuration terminates successfully, then the other one must terminate successfully as well.
-The fact that the observation relation is split into two parts means that we really have a logical approximation where one machine approximates the behavior of the other machine.
+\]
 

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Lars comments.

--- 2018-07-09 fully-abstract-comp/popl2019/paper.tex
@@
 
+\todo[inline]{In the below, Lars would like to be more precise about the ``in a general way'' below.
+  e.g., by contrasting it to our earlier work where the WBCF and LSE properties were only proved via specific examples.}
 This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
@@
 % Return pointers/return seals
-The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
+The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
 Both of the capabilities in the pair are of course sealed with the same seal.
@@
 % Note about them vs us, we do nothing that they couldn't do
-In the description of the StackToken calling scheme, it has sometimes been in terms of ``them vs us''.
+In the description of the \stktokens{} calling scheme, it has sometimes been in terms of ``them vs us''.
 This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
@@
 The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
-When we have defined the overlay semantics, we prove full-abstractness for a compilation from the overlay semantics to the underlying semantics.
+When we have defined the overlay semantics, we prove full-abstraction for a compilation from the overlay semantics to the underlying semantics.
 Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
typos and removed todos

@@
 
-\todo[inline]{Describe $H^\mathrm{code,\square}_\sigma$}
-
-the seal interpretation function $h^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
+The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
 the return seals can only be used to seal return pointers, but only when they actually represent meaning full returns.
@@
 \subsubsection{Value relations}
-  \todo[inline]{Do we cut the sealed case of the untrusted value relation?}
+The purpose of the value relations is two fold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
 

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
More work on section 5.

@@
 % Further, all stack frames should be in addresses above $\stkb$.
+The seal interpretation functions are in the pure regions, so the heap relation ($\lrheap$) also makes sure that only one seal interpretation function has an interpretation for each seal.
 \extend{Write about the particulars of the three memory satisfaction: heap relation makes sure only one interpretation for each seal. Free stack $\stkb$}
@@
 The purpose of the value relations is two fold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
+As previously explained, we try to ensure LSE and WBCF for some trusted code of which we require certain things like treating return seals in a reasonable way.
+This means that it is safe to give return seals to trusted code.
+At the same time, we have some code that we do not trust, and given a return seal it may try to abuse it to break LSE and WBCF for the trusted code.
+This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and untrusted value relation.
 
@@
 %
-    \left\{
-    \begin{array}{l}
-      \npair{\left(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T} \right)} \; | \; \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
-      \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
-      \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\ \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
-      % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-    \end{array}
-    \right\}\cup\\
-    \left\{
-    \begin{array}{l}
-    \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)}\; | \; [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
-      \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
-    \end{array}
-    \right\} \cup \\
     \left\{
@@
     \end{array}
-    \right\} \cup \\ \dots
+    \right\} \cup \\
+%
+    \left\{
+    \begin{array}{l}
+    \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)}\; | \; [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
+      \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+    \end{array}
+    \right\} \cup \\
+        \left\{
+    \begin{array}{l}
+      \npair{\left(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T} \right)} \; | \; \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
+      \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
+      \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
+      \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \tand \dots
+      % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
+      % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+    \end{array}
+    \right\}\cup \dots
 %
@@
 \end{multline*}
-
+A stack pointers is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
+In addition depending on the permission, a read and write condition needs to be satisfied.
+Intuitively, the read condition guarantees that the capability can only be used to read safe values, i.e.\ values in $\mathcal{V}_\untrusted$.
+Similarly the write condition says that the capabilities can at least be used to write the safe values.
+These two properties are expressed by picking out the regions from the world that govern the memory the stack pointer gives access to.
+The free stack is modelled by the free stack world, so the regions must come from that part of the world.
+Further, stack pointers are linear, so the regions they depend on must be addressable which means that they must be owned by that free stack world.
+% The write condition is called $\stackWriteCond{}$
+% \begin{multline*}
+%   \stackWriteCond{W} =\\
+%   \left\{
+%     \begin{array}{l}\npair{A} \mid 
+%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
+%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
+%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
+%     \end{array}
+%   \right\}
+% \end{multline*}
+
+The untrusted value relation also relates sets of seals that are syntactically identical and that are disjoint from the return and closure seals of interest ($\gsigrets$ and $\gsigcloss$). 
+Further, the seals should only be used to seal words from the untrusted value relation, so the seal interpretation function for each of the seals must be the untrusted value relation.
+
+The untrusted value relation also needs to include sealed capabilities in order to allow return pairs and closures to be given to untrusted code.
+The sealed capabilities should of course be sealed with the same seal, and the seal interpretation function for that seal should be equal to the one of the code region.
+Aside from this, the pair of sealables sealed by the seal should satisfy two things: (1) they should be in the seal interpretation function for that seal and (2) in any possible future\footnote{The $\purePart{}$ function strips away all ownership from a world.
+A non-linear capability should not depend on linear things, so we take any future of $\purePart{W}$.}, it should be safe to jump to the sealable pair when they are paired with sealables permitted by the seal interpretation function for that seal, so the quadruple must be in the $\mathcal{E}_\mathrm{xjmp}$ relation.
+
+We have left out the part of the untrusted value relation that relates memory capabilities and integers (they are related when they are equal).
+The conditions for related memory capabilities are almost the same as for stack pointers.
+The main difference is that the conditions depend on the heap part of the world rather than the free stack.
+Further, memory capabilities can be executable which needs extra care to ensure security which we will see in the trusted value relation.
+
+The trusted value relation includes everything that the untrusted value relation contains and a few extra things.
 \begin{multline*}
@@
     \; | \; \exists r \in \dom(\pwheap) \ldotp \\
-              \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc} \tand \dom(\code) \subseteq \ta \\
+              \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\mscode,\gc} \tand \dom(\mscode) \subseteq \ta \\
               \quad \tand [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \tand \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss
@@
 \end{multline*}
-
-\begin{multline*}
-  \stackWriteCond{W} =\\
-  \left\{
-    \begin{array}{l}\npair{A} \mid 
-      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
-      \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
-      \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
-    \end{array}
-  \right\}
-\end{multline*}
+A side from the sets of seals in the untrusted value relation, sets of return and closure seals are in the trusted value relation when they belong to some trusted code.
+This means that in the heap part of the world there must be a code region that has the seals in question, and the code should be in $\ta$.
+
+The trusted value relation also contains executable capabilities for the trusted code.
+Generally for executable code capabilities, the $\xReadCond{}$ should hold which requires there to be a code region that governs the piece of code memory that the capability has authority over.
 \[
@@
       \exists r \in \addressable{\normal, \pwheap} \ldotp \\
-       \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\tand
-       \dom(\code) \supseteq A 
+       \pwheap(r) \nequal \codereg{\_,\_,\mscode,\gc}\tand
+       \dom(\mscode) \supseteq A 
     \end{array}
@@
 
+With the logical relation defined, we are ready to state the fundamental theorem of logical relations (FTLR).
 \begin{lemma}[FTLR]
@@
 \end{lemma}
+It roughly says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
+This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ ensures, and for trusted code the semantic reasonability condition must also be satisfied by the code.
+The proof of the FTLR is done by a nested induction that categorises and considers each possible step the two executions can take.
+While the lemma is not used directly in the proof sketch we present in Section~\ref{subsec:proof-sketch}, it is detrimental to the proof of the lemmas we use in it.
 
 \subsection{Component logical relation}
-
-\begin{multline*}
-  \lrcomp(W) =\\
-  \left\{\begin{aligned}
-      &\npair{\var{comp},\var{comp}} \;\mid \;\\
-      &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
-      &\qquad\text{For all } W' \future W \ldotp \\
-      &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
-      &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
-      &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
-    \end{aligned}
-  \right\}\\
-\cup \left\{
-    \begin{multlined}
-      \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
-      \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
-      \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
-    \end{multlined}
-  \right\} 
-\end{multline*}
+The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
+The component relation relates a component $(\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss)$ to itself when two conditions are satisfied.
+First, when words that relate to them selves in the untrusted value relation are used to satisfy the imports, i.e.\ the words are placed on the import addresses in $\msdata$, and this data memory is combined with the code memory $\mscode$, then it forms a safe heap, i.e.\ it is in the $\mathcal{H}$ relation.
+Second, the exports should always be safe to use which means that they must be in the untrusted value relation in any future world.
+
+% \begin{multline*}
+%   \lrcomp(W) =\\
+%   \left\{\begin{aligned}
+%       &\npair{\var{comp},\var{comp}} \;\mid \;\\
+%       &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+%       &\qquad\text{For all } W' \future W \ldotp \\
+%       &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+%       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
+%       &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
+%       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+%     \end{aligned}
+%   \right\}\\
+% \cup \left\{
+%     \begin{multlined}
+%       \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
+%       \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
+%       \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
+%     \end{multlined}
+%   \right\} 
+% \end{multline*}
 
@@
 \subsection{Proof sketch}
+\label{subsec:proof-sketch}
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Rephrase the introduction to the overlay semantics and reasonability stuff.

--- 2018-07-09 fully-abstract-comp/popl2019/paper.tex
@@
 
-\todo[inline]{In the below, Lars would like to be more precise about the ``in a general way'' below.
-  e.g., by contrasting it to our earlier work where the WBCF and LSE properties were only proved via specific examples.}
 This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
@@
 % Return pointers/return seals
-The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
+The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
 Both of the capabilities in the pair are of course sealed with the same seal.
@@
 % Note about them vs us, we do nothing that they couldn't do
-In the description of the \stktokens{} calling scheme, it has sometimes been in terms of ``them vs us''.
+In the description of the StackToken calling scheme, it has sometimes been in terms of ``them vs us''.
 This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
-However, \stktokens{} is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
+However, \stktokens{} is a modular calling scheme: no restriction is put on adversarial components that we do not expect trusted components to meet.
 This means that any component can ensure WBCF and LSE by employing \stktokens{}.
@@
 Our proof uses a novel approach we call fully abstract overlay semantics.
-In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
-An overlay semantics is basically a different view on an existing semantics.
-For instance, it may have new syntactic constructs and have special semantics for these things, but they correspond to something on the underlying machine.
-In the overlay semantics we are about to define for \trgcm{}, an explicit stack pointer construct is part of the new view.
-The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
-When we have defined the overlay semantics, we prove full-abstraction for a compilation from the overlay semantics to the underlying semantics.
-Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.
-In our case, the compilation even turns out to be the identity function!
-
-% protection against others purpose of cc
-We can not show that \stktokens{} guarantee much if the security mechanisms of the capability machine are undermined by a poorly configured machine, e.g.\ the same seal is present in different components or the stack capability have aliases.
-It is also possible to for a caller to undermine the security mechanisms during execution.
-For example, the caller could give the return seal for a call to the callee.
-\stktokens{} do not protect a component against itself as the component is assumed to behave reasonably.
-In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components from unsafe behavior.
+The idea is to define a second operational semantics for programs in our target language.
+This second semantics uses a different abstract machine and different run-time values, but it executes in lock-step with the original semantics and there is a very close correspondence between the state of both machines.
+
+The main difference between the two semantics, is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
+Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
+For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.
+
+The fact that \stktokens{} enforces LSE and WBCF is formulated as the theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
+The theorem states that this function consitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
+Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than the corresponding components in the well-behaved overlay semantics.
+In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
+More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics iff they are indistinguishable to arbitrary other components in the overlay semantics.
+
+Our formal results are complicated by the fact that they only hold for well-behaved components that respect the basic rules of the calling convention, and also on a sane initial configuration of the system.
+For example, the system should be setup such that seals that components use for constructing return pointers should not be shared with other components.
+We envision distributing seals as a job for the linker, so this means our results depend on a linker to do this properly.
+As another example, a seal that is used to construct a return pointer, can be used again to construct other return pointers for the same return point, but a different seal must be used for others.
+Such seals should also never be passed to other components.
+Such requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
+The general pattern is that $\stktokens{}$ only protects components that do not shoot themselves in the foot by violating a few basic rules.
+In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components to do certain unsafe things.
 
@@
     \hline
-    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
+    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{(\perm,\baddr,\eaddr,\aaddr} $ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
     \hline
@@
 
-The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
+\todo[inline]{Describe $H^\mathrm{code,\square}_\sigma$}
+
+the seal interpretation function $h^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
 the return seals can only be used to seal return pointers, but only when they actually represent meaning full returns.
@@
 \subsubsection{Value relations}
-The purpose of the value relations is two fold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
+  \todo[inline]{Do we cut the sealed case of the untrusted value relation?}
 

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Another slight reformulation.

@@
 % Back to high-level view
-In \stktokens{} the base address of the memory $\stkb$ used for the stack is agreed upon upfront.
-As \srccm{}'s view builds-in the stack, it needs to know the base address during execution.
-From the perspective of a well-behaved component, other components can do whatever they want for their calls and local state as long as they cannot break WBCF and LSE for the well-behaved component.
-Due to this, \srccm{} has a set of trusted addresses $\ta$ that are allowed to perform machine calls that push stack frames to the call stack.
-To make these constants available during execution, they are passed around in the \trgcm{} step relation.
+As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
+However, there is no such assumption on the untrusted other components they interact with.
+To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
+Only instructions at these addresses are allowed to perform calls that push stack frames to the call stack, i.e.\ calls for which \srccm{} provides native LSE and WBCF guarantees.
+The constant $\ta$ is passed around as a parameter of the \trgcm{} step relation.
+Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
+
 Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Typo.

@@
 
-By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack which means that it insures LSE.
+By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack which means that it ensures LSE.
 Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack which means that we have WBCF.

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Shorten a phrase.

@@
 
-By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack which means that it ensures LSE.
-Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack which means that we have WBCF.
-In other words by inspection of the operational semantics, \srccm{} has LSE and WBCF.
+By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack.
+Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack.
+In other words by inspection of the operational semantics, \srccm{} guarantees LSE and WBCF.
 It is important to point this out as \stktokens{} guarantee exactly the LSE and WBCF we have on \srccm{}.

### 2018-07-09 fully-abstract-comp/technical-report.tex
TR typos
typo
TR todo
TR TODO and quick formatting issue fix

@@
 Memory satisfaction for new worlds:
+\lau{09-07-2018: TODO: add $\gc$ to memsat relations.}
 \[
@@
 
+\lau{09-07-2018: Fix these definitions' formatting like in paper} 
 \[
@@
       \forall i \leq n \ldotp \\
-      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \Rightarrow (\ms_T,\reg_T) \term\\
+      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
     \end{array}
@@
       \forall i \leq n \ldotp \\ 
-      \quad (\ms_T ,\reg_T) \term[i] \Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
+      \quad (\ms_T ,\reg_T) \term[i] \\\qquad\Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
     \end{array}
@@
       \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
-      \quad \exists R : S \fun \powerset{\nats} \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
       \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
@@
       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
-      \quad \exists R : S \fun \powerset{\nats} \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
       \qquad \forall r \in S \ldotp |R(r)| = 1\\
@@
   \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
-H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; \gc,
-H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc,
+H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
 \]
@@
       \forall a \in \dom(\code)\ldotp\\
-      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})}) \wedge
+      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
     \end{array}
@@
 
+\lau{09-07-2018: TODO: It seems very odd to me that the return and closure seals in $\gc$ are independent of the seals in the component.
+I would suspect that in the trusted case, $\gsigrets,\gsigcloss$ is a suberset of the seals in the component (or something like that).
+}
 \begin{lemma}[FTLR for components]

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
First (rough) draft of section 5 done.

--- 2018-07-09 fully-abstract-comp/popl2019/paper.tex
@@
 Our proof uses a novel approach we call fully abstract overlay semantics.
-The idea is to define a second operational semantics for programs in our target language.
-This second semantics uses a different abstract machine and different run-time values, but it executes in lock-step with the original semantics and there is a very close correspondence between the state of both machines.
-
-The main difference between the two semantics, is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
-Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
-For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.
-
-The fact that \stktokens{} enforces LSE and WBCF is formulated as the theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
-The theorem states that this function consitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
-Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than the corresponding components in the well-behaved overlay semantics.
-In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
-More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics iff they are indistinguishable to arbitrary other components in the overlay semantics.
-
-Our formal results are complicated by the fact that they only hold for well-behaved components that respect the basic rules of the calling convention, and also on a sane initial configuration of the system.
-For example, the system should be setup such that seals that components use for constructing return pointers should not be shared with other components.
-We envision distributing seals as a job for the linker, so this means our results depend on a linker to do this properly.
-As another example, a seal that is used to construct a return pointer, can be used again to construct other return pointers for the same return point, but a different seal must be used for others.
-Such seals should also never be passed to other components.
-Such requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
-The general pattern is that $\stktokens{}$ only protects components that do not shoot themselves in the foot by violating a few basic rules.
-In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components to do certain unsafe things.
+In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
+An overlay semantics is basically a different view on an existing semantics.
+For instance, it may have new syntactic constructs and have special semantics for these things, but they correspond to something on the underlying machine.
+In the overlay semantics we are about to define for \trgcm{}, an explicit stack pointer construct is part of the new view.
+The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
+When we have defined the overlay semantics, we prove full-abstraction for a compilation from the overlay semantics to the underlying semantics.
+Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.
+In our case, the compilation even turns out to be the identity function!
+
+% protection against others purpose of cc
+We can not show that \stktokens{} guarantee much if the security mechanisms of the capability machine are undermined by a poorly configured machine, e.g.\ the same seal is present in different components or the stack capability have aliases.
+It is also possible to for a caller to undermine the security mechanisms during execution.
+For example, the caller could give the return seal for a call to the callee.
+\stktokens{} do not protect a component against itself as the component is assumed to behave reasonably.
+In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components from unsafe behavior.
 
@@
 % Back to high-level view
-As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
-However, there is no such assumption on the untrusted other components they interact with.
-To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
-Only instructions at these addresses are allowed to perform calls that push stack frames to the call stack, i.e.\ calls for which \srccm{} provides native LSE and WBCF guarantees.
-The constant $\ta$ is passed around as a parameter of the \trgcm{} step relation.
-Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
-
+In \stktokens{} the base address of the memory $\stkb$ used for the stack is agreed upon upfront.
+As \srccm{}'s view builds-in the stack, it needs to know the base address during execution.
+From the perspective of a well-behaved component, other components can do whatever they want for their calls and local state as long as they cannot break WBCF and LSE for the well-behaved component.
+Due to this, \srccm{} has a set of trusted addresses $\ta$ that are allowed to perform machine calls that push stack frames to the call stack.
+To make these constants available during execution, they are passed around in the \trgcm{} step relation.
 Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
@@
 \begin{definition}[Plugging a component into a context]
+  \label{def:comp-context-plugging}
   When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
@@
 % \end{itemize}
-\todo[inline]{Describe figure here or save it for proof sketch subsection?}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
-    % \draw[help lines,yellow] (0,0) grid (10,7);
-    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
-
-    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
-    \node at (5,4) { $\mathrel{\Rightarrow}$ };
-    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
-
-    \node at (4.35,2.8) { (1) };
-    \node at (5,3.6) { (2) };
-    \node at (5.65,2.8) { (3) };
-
-    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
-
-    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
-
-    \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
-    \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
-    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
-    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
-    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
-
-    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
-    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
-    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
-
-    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
-
-    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
-  \end{tikzpicture}
-  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
-\end{figure}
 
 \subsection{Logical relation}
-\label{subsec:worlds}
+\label{subsec:logical-relation}
 The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
@@
 \subsubsection{Worlds}
+\label{subsec:worlds}
 Generally speaking, worlds are used in a logical relation to model the memory of a language.
@@
 
-\subsection{Component logical relation}
-The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
-The component relation relates a component $(\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss)$ to itself when two conditions are satisfied.
-First, when words that relate to them selves in the untrusted value relation are used to satisfy the imports, i.e.\ the words are placed on the import addresses in $\msdata$, and this data memory is combined with the code memory $\mscode$, then it forms a safe heap, i.e.\ it is in the $\mathcal{H}$ relation.
-Second, the exports should always be safe to use which means that they must be in the untrusted value relation in any future world.
-
+\subsection{Full abstraction proof sketch}
+\label{subsec:proof-sketch}
+The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward.
+The component relation relates components to them self which makes sense as well-formed components should contain no \srccm{} specific constructs.
+We prove a number of lemmas about the relations.
+The first is a FTLR for components that says (1) all well-formed untrusted components are related to themselves in the component relation, and (2)  all well-formed and reasonable trusted components are related to themselves in the component relation.
+% \subsection{Component logical relation}
+% The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
+% The component relation relates a component $(\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss)$ to itself when two conditions are satisfied.
+% First, when words that relate to them selves in the untrusted value relation are used to satisfy the imports, i.e.\ the words are placed on the import addresses in $\msdata$, and this data memory is combined with the code memory $\mscode$, then it forms a safe heap, i.e.\ it is in the $\mathcal{H}$ relation.
+% Second, the exports should always be safe to use which means that they must be in the untrusted value relation in any future world.
 % \begin{multline*}
@@
 % \end{multline*}
-
 \begin{lemma}[FTLR for components]
@@
   \begin{itemize}
-  \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
   \item $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$
@@
     \begin{itemize}
-    \item $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component (see Section~\ref{sec:reasonability})
+    \item $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component
     \item $\dom(\comp.\mscode) \mathrel{\#} \ta$
@@
   Then there exists a $W$ such that
-  \begin{itemize}
-  \item $\npair{(\comp,\comp)} \in \lrcomp(W)$
-  \item $\dom(\pwheap)$ can be chosen to not include any finite set of region names 
-  \item $\dom(\pwfree) =
-    \dom(\pwpriv)=\emptyset$
-  \end{itemize}
+  $\npair{(\comp,\comp)} \in \lrcomp(W)$
 \end{lemma}
+The next is a compatibility lemma for plugging that says that given a context and component each related to itself in the component relation, then plugging the context with the component will give an executable configuration in the executable configuration relation.
 \begin{lemma}[Compatibility lemma for context plugging]
@@
 \end{lemma}
-
-\subsection{Executable configuration logical relation}
-
-\begin{multline*}
-  \lrec[\square,\gc = (\ta,\stkb)](W) = \\
-\left\{
-  \begin{array}{l}
-     \npair{\left(
-    (\ms_S,\reg_S,\stk,\ms_\stk),
-    (\ms_T,\reg_T)\right)} \mid \\
-    \quad \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
-    \qquad \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
-    \qquad \reg_S(\pcreg) \neq \retptrc(\_) \tand
-    \reg_S(\rdata) \neq \retptrd(\_) \tand \\
-    \qquad \nonExec{\reg_S(\rdata)} \tand  
-     \nonExec{\reg_T(\rdata)} \tand \\
-    \qquad \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
-  \end{array}
-\right\}
-\end{multline*}
-\begin{lemma}[Adequacy of execution configuration LR]
+Finally, we have an adequacy lemma for the execution configuration relation.
+This lemma says that if an \srccm{} configuration approximates a \trgcm{} configuration, and the \trgcm{} configuration terminates, then so does the \srccm{} configuration.
+The symmetric case is also true.
+% \begin{multline*}
+%   \lrec[\square,\gc = (\ta,\stkb)](W) = \\
+% \left\{
+%   \begin{array}{l}
+%      \npair{\left(
+%     (\ms_S,\reg_S,\stk,\ms_\stk),
+%     (\ms_T,\reg_T)\right)} \mid \\
+%     \quad \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
+%     \qquad \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
+%     \qquad \reg_S(\pcreg) \neq \retptrc(\_) \tand
+%     \reg_S(\rdata) \neq \retptrd(\_) \tand \\
+%     \qquad \nonExec{\reg_S(\rdata)} \tand  
+%      \nonExec{\reg_T(\rdata)} \tand \\
+%     \qquad \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
+%   \end{array}
+% \right\}
+% \end{multline*}
+\begin{lemma}[Adequacy of execution configuration logical relation]
   \label{lem:adequacy}
@@
 \end{lemma}
-
-\subsection{Proof sketch}
-\label{subsec:proof-sketch}
+We use these three lemmas to prove Theorem~\ref{thm:full-abstraction}
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  \todo[inline]{Rewrite}
+  The proof of both directions are very similar, so we only show the right direction.
+  W.l.o.g.\ it suffices to show one direction of the contextual equivalence, so we assume ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
+  The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.
+  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that $\comp_1$ and $\comp_2$ are well-formed and reasonable.
+  Further, by $\tconeq$ the context $\context$ must be well-formed.
+  We prove arrow (1) in the figure by using the mentioned assumption about $\comp_1$ and $\context$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
+  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$, so by the assumption that $\comp_1$ and $\comp_2$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ which proves arrow (2) in the figure.
+  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use that $\comp_2$ is well-formed and reasonable and that $\context$ is well-formed.
+\end{proof}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
+    % \draw[help lines,yellow] (0,0) grid (10,7);
+    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
+
+    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
+    \node at (5,4) { $\mathrel{\Rightarrow}$ };
+    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
+
+    \node at (4.35,2.8) { (1) };
+    \node at (5,3.6) { (2) };
+    \node at (5.65,2.8) { (3) };
+
+    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
+
+    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
+
+    \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
+    \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
+    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
+    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
+    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
+
+    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
+    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
+    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
+
+    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
+
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
+  \end{tikzpicture}
+  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
+  \label{fig:fa-proof-sketch}
+\end{figure}
+
+% \subsection{Proof sketch}
+% \label{subsec:proof-sketch}
+% \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
   % \item Consider first the upward arrow.
@@
 
-  First consider the right arrow:
-
-    Assume $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$.
+%   First consider the right arrow:
+
+%     Assume $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$.
 Take $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$.
-%
-    Take a $\trg{\context}$ such that $\vdash \trg{\context}$ and we will prove that
-    $\trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow
-    \trg{\plug{\context}{\var{comp}_2} \term}$.
-%
-    By symmetry, we can assume w.l.o.g. that $\trg{\plug{\context}{\var{comp}_1} \term}$ and prove that $\trg{\plug{\context}{\var{comp}_2} \term}$.
-    Note that this implies that $\trg{\context}$ is a valid context for both $\trg{\var{comp}_1}$ and $\trg{\var{comp}_2}$.
-%
-    First, we show that also $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
-    Take $n$ the amount of steps in the termination of $\plug{\context}{\var{comp}_1} \trg{\term}$.
-    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
-    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
-    Together with $\plug{\context}{\var{comp}_1} \trg{\term[n]}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
-%
-    It follows from $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
-%
-    It now remains to show that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
-    Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
-    It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
-    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$
-    Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_2} \trg{\term}$, concluding the second direction of the proof.
-
- The left arrow is proven in a similar manner.
-\end{proof}
+% %
+%     Take a $\trg{\context}$ such that $\vdash \trg{\context}$ and we will prove that
+%     $\trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow
+%     \trg{\plug{\context}{\var{comp}_2} \term}$.
+% %
+%     By symmetry, we can assume w.l.o.g. that $\trg{\plug{\context}{\var{comp}_1} \term}$ and prove that $\trg{\plug{\context}{\var{comp}_2} \term}$.
+%     Note that this implies that $\trg{\context}$ is a valid context for both $\trg{\var{comp}_1}$ and $\trg{\var{comp}_2}$.
+% %
+%     First, we show that also $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
+%     Take $n$ the amount of steps in the termination of $\plug{\context}{\var{comp}_1} \trg{\term}$.
+%     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+%     It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+%     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
+%     Together with $\plug{\context}{\var{comp}_1} \trg{\term[n]}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
+% %
+%     It follows from $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
+% %
+%     It now remains to show that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
+%     Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
+%     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
+%     It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+%     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$
+%     Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_2} \trg{\term}$, concluding the second direction of the proof.
+
+%  The left arrow is proven in a similar manner.
+% \end{proof}
 

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex

@@
 
-By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack which means that it insures LSE.
-Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack which means that we have WBCF.
-In other words by inspection of the operational semantics, \srccm{} has LSE and WBCF.
+By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack.
+Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack.
+In other words by inspection of the operational semantics, \srccm{} guarantees LSE and WBCF.
 It is important to point this out as \stktokens{} guarantee exactly the LSE and WBCF we have on \srccm{}.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Typo
Some shortening.

@@
 \]
-  \caption{The syntax of our capbility machine with seals and linear capbilities.}
+  \caption{The syntax of our capability machine with seals and linear capbilities.}
   \label{fig:target-syntax}
@@
 % Cap machine inspired by CHERI, but with linear capabilities
-In this section, we introduce a simple but representative capability machine with linear capabilities.
-We will later introduce a capability machine similar to the one in this section, so to easily distinguish the two, we call this one \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main exception.
-In order to avoid details uninteresting to the focus of this work, \trgcm{} assumes an infinite address space and unbounded integers.
-\lau{Do we have a good reason to assume an infinite address space for this work?
-  We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
+In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
+For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
+% \lau{Do we have a good reason to assume an infinite address space for this work?
+%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
 
@@
 Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
-\trgcm{} only has memory capabilities, and we limit them to a number of sensible permissions making the available permissions: read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$).
-We define the preorder $\le$ on permissions as illustrated in Figure~\ref{fig:perm-hier}.
+The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
 % Pointer and linearity
-In addition to the two basic things, the capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
-The linearity of a capability can either be $\linear$ or $\normal$ where $\normal$ corresponds to a traditional capability.
+In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
+The linearity $\lin$ is either $\normal$ for traditional capabilities $\linear$ for linear ones.
 % What is linear capabilities, split/splice
 A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or stored to memory, the source register is cleared, and when a linear capability is read from memory, the source address is cleared.
-We will see this later in this section when we get to the operational semantics.
-The non-duplicability of linear capabilities means that a linear capability cannot suddenly be aliased if it wasn't to begin with.
+This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source register or memory location is cleared.
+The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten.

@@
 
-\begin{wrapfigure}{r}{0.33\linewidth}
+\begin{wrapfigure}{r}{0.28\linewidth}
   \centering
@@
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have sealed capabilities to achieve this.
-Syntactically, sealed capabilities add two things to the capability machine: Sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ and sealed capabilities $\sealed{\sigma,\vsc}$.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
+Syntactically, sealed capabilities add two types of capabilities.
 % seals
-A set of seals consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$ which is represented by a natural number.
+A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
 % sealed caps
-A sealed capability consists of a sealable $\vsc$ and the seal $\sigma$ it has been sealed with.
-The sealables are the sets of seals and the capabilities, and we refer to the set of them as $\SealableCaps$.
-The system native tokens $\Cap$ on \trgcm{} are the sealables and the sealed capabilities.
+A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
+Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
 
 %%% Domains
-The words on \trgcm{} are the system native tokens in $\Cap$ and data which is represented by $\ints$.
-We assume a finite set of register names $\RegName$ that at least contains register names $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
-We have a notion of a complete memory where all memory addresses contains a word and a memory segment that need not contain words on all addresses.
-Both are represented as functions from addresses, defined as the natural numbers, to words, but the memory segment functions only need to be partial.
-$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.\ an attempt to break the capability machine guarantees.
+The words on \trgcm{} are capabilities and data (represented by $\ints$).
+We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
+We have complete memories mapping all addresses to words and memory segments mapping some addresses to words (i.e.\ partial functions).
+$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.\ an out-of-bounds memory access.
 An executable configuration is a register file and memory pair.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten.

@@
 %%% Domains
-The words on \trgcm{} are capabilities and data (represented by $\ints$).
+Words on \trgcm{} are capabilities and data (represented by $\ints$).
 We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
@@
 % syntax (instructions)
-\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability machine specific instructions.
-The instructions that can be considered standard are: unconditional and a conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability machine specific instructions simply inspect the type of system native tokens or project one of their fields.
-These are \texttt{gettype} (for getting the type of a system token) and \texttt{getp}, \texttt{getl}, \texttt{getb}, \texttt{gete}, and \texttt{geta} (for projecting permission, linearity, base, end, and current address(/seal), respectively).
-There are two instructions for manipulating the current address(/seal) of a capability(/set of seals): \texttt{cca} changes the current address(/seal) of a capability(/set of seals) by some offset, and \texttt{seta2b} sets the current address(/seal) of a capability(/set of seals) to the base address(/seal).
+\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability-related instructions.
+The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
+The simplest of the capability instructions simply inspect the properties of capabilities: type (\textt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
+The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
 The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
@@
 Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two rather than reducing the range of authority (\texttt{split}).
-\trgcm{} also has an instruction that does the inverse of splitting, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
-\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities by crossing security boundaries (\texttt{xjmp}, see below).
+The reverse is possible using \texttt{splice}.
+Sealables can be sealed using \texttt{cseal} and pairs of sealed capabilities can be unsealed by crossing security boundaries (\texttt{xjmp}, see below).
 Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten.

@@
 For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In \stktokens{}, the caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to the base address of the return token they gave to the callee.
+In \stktokens{}, the base address of the stack is fixed as a compile-time constant.
+The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, as it was for the return token they gave to the callee.
 In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
@@
 % KJAA: It is unclear to me what 'at this point' refers to.
-At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
-However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
-In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
-To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
-The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
-The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
-Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
-At this point, the adversary has everything they need to return from either of the two callbacks.
-Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
-From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
-They are both correct as the adversarial component has created two call stacks, one for each of the components.
-In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
-We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
-A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
-This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
-
-Having a globally decided stack address also solves another challenge.
-Namely, the stack capability has to be linear in order for a caller to use it as a token, so a caller has to make sure that a capability is in fact linear when it is proposed as the stack capability by an untrusted source.
-This is, however, implicitly checked when the caller checks that the stack capability governs the global stack base address.
-The reason is that the stack capability needs to start execution as a linear capability which means that the capability that governs the global stack base address to begin with is linear.
-Due to the non-duplicability of linear capability and the fact that they cannot become non-linear during execution, the caller knows that if they have a capability that governs the global stack base address, then it must be linear.
+% At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
+% However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
+% In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
+% To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
+% The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
+% The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
+% Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
+% At this point, the adversary has everything they need to return from either of the two callbacks.
+% Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
+% From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
+% They are both correct as the adversarial component has created two call stacks, one for each of the components.
+% In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
+% We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
+% A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
+% This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
+
+In \stktokens{} we ensure that at the start of execution, the stack memory is only referenced by a single, linear stack capability.
+Because of this, we can assert validity of the return token simply by checking its base address and splicing it with the caller's stack frame.
+There is no need to check linearity, because we know that only linear capabilities to this memory exist.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Fix typesetting problem

@@
 Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
-If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than $\tload$'s worst-case...
+If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst-case...
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Latex fixes.

@@
 Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
-As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like $\tmove{}$, $\tstore{}$ and $\tload{}$.
+As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
 Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
-Our instructions are a bit clumsy from this point-of-view because, for example, $\tmove{}$ or $\tstore{}$ will zero the source register resp. memory location if the value being written is linear.
+Our instructions are a bit clumsy from this point-of-view because, for example, \texttt{move} or \texttt{store} will zero the source register resp.
+memory location if the value being written is linear.
 A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
@@
 
-Another problem is caused by the move semantics for $\tload{}$ in the presence of multiple hardware threads.
+Another problem is caused by the move semantics for \texttt{load} in the presence of multiple hardware threads.
 In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a $\tload{}$ of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
-This is in principle not a problem except that a compare-and-swap is significantly slower than a regular $\tload{}$ (on the order of 10x slower or more).
+This means that a \texttt{load} of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
+This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
 When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten, fix typos.

@@
 In the first one, we use the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
-When a word $w$ that can possibly be a linear capability is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
+When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
 In the case of \texttt{move}, the source register $r_2$ is overwritten with $\linCons{\Phi(r_2)}$.
@@
 Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
-Clearing a memory address corresponds to writing $0$ to the memory, but in order to write to memory we need to provide a capability with write-permission.
-This means that to load a linear capability from memory, we have to use a capability with read and write permission.
+In this case, we require that the memory capability used for loading also has write-permission.
 
 % geta, cca
-The instructions \texttt{geta} and \texttt{cca} are fairly simple: \texttt{geta} projects the current address or current seal from a capability or a set of seals, respectively, and it returns the error value $-1$ for data and sealed capabilities.
-The \texttt{cca} instruction changes the current address or seal of a capability or set of seals, respectively, by a given offset.
+The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
+\texttt{cca} changes the current address or seal of a capability or set of seals, respectively, by a given offset.
 It handles linearity by updating the capability in-place, i.e.\ the source register is used as the target register.
 This is the second approach we use to uniformly handle linearity.
-The \texttt{jmp} instruction is a fairly simple jump that just sets the $\pcreg$-register.
+The \texttt{jmp} instruction is a simple jump that just sets the $\pcreg$ register.
 
@@
 % xjmp
-Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities unseals them and put one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the capability placed in $\rdata$) is non-executable.
-A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment of the closure.
+Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
+A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
 Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
-This also makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
-The opaque sealed capabilities encapsulate a security domain's local state and authority, and it only becomes accessible again when control is transferred to the security domain.
-Some care should be taken when sealing as reuse of the same seal for multiple closures would make it possible to jump to the code of one capability with the environment of another.
+This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
+Opaque sealed capabilities encapsulate a security domain's local state and authority, and it only becomes accessible again when control is transferred to the security domain.
+Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one capability with the environment of another.
 % xjmp to unseal
-\trgcm{} does not have an instruction for unsealing single sealed capabilities directly (using the seal that was used to do the sealing), but it can be (partially) simulated using \texttt{xjmp}.
+\trgcm{} does not have an instruction for unsealing capabilities directly, but it can be (partially) simulated using \texttt{xjmp}.
 
@@
 With \texttt{split}, we can reduce the range of authority of a linear capability without losing any authority as we retain it in the second capability.
-However, if we only have \texttt{split}, then in some sense splitting a capability is not completely lossless as we may have authority over the same range of addresses after a \texttt{split}, but we have that authority in two capabilities rather than just one.
-In order to make up for this, we have a \texttt{splice} instruction that essentially does the inverse of \texttt{split}.
-Namely given two capabilities with adjacent ranges of authority and the same permissions and linearity, it splices them together into one capability.
+The \texttt{splice} instruction essentially does the inverse of \texttt{split}.
+Given two capabilities with adjacent ranges of authority and the same permissions and linearity, \texttt{splice} splices them together into one capability.
 The two instructions work in the same way for seal sets.
@@
 In Section~\ref{sec:form-secur-with}, we will get back to how a program is instrumented to create an executable configuration.
-We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
@@
 \]
-\vspace{0.75cm}
+\vspace{0.5cm}
 \begin{mathpar}
@@
 \label{sec:stktokens-explained}
-In this section, we introduce our calling convention \stktokens{} that insures LSE and WBCF.
+In this section, we introduce our calling convention \stktokens{} that ensures LSE and WBCF.
 We will gradually explain each of the security measures \stktokens{} take and motivate them with the attacks they prevent.
@@
 % Call/return
-The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem}, and it may look daunting due to a fair bit of bookkeeping, but it essentially does the following: Given a stack pointer in the stack register, $42$ is pushed to the stack to insure that the stack frame will be non-empty.
+The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem}, and it may look daunting due to a fair bit of bookkeeping, but it essentially does the following: Given a stack pointer in the stack register, $42$ is pushed to the stack to ensure that the stack frame will be non-empty.
 Based on the stack pointer in $\rstk$, a stack pointer for the callee is created (this corresponds to the return token of \stktokens{}) as well as a return pointer which governs the private part of the stack, i.e.\ addresses $[\aaddr_\stk,\eaddr_\stk]$ which is $42$ and everything (conceptually) below on the stack.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Win a bit of space

@@
 
-    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\Small Contextual equivalence preservation} (0,1);
   \end{tikzpicture}

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
More work on section 5 and notes for discussion.

--- 2018-07-10 fully-abstract-comp/popl2019/paper.tex
@@
 % syntax (instructions)
-\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability-related instructions.
-The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability instructions simply inspect the properties of capabilities: type (\textt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
-The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
+\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability machine specific instructions.
+The instructions that can be considered standard are: unconditional and a conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
+The simplest of the capability machine specific instructions simply inspect the type of system native tokens or project one of their fields.
+These are \texttt{gettype} (for getting the type of a system token) and \texttt{getp}, \texttt{getl}, \texttt{getb}, \texttt{gete}, and \texttt{geta} (for projecting permission, linearity, base, end, and current address(/seal), respectively).
+There are two instructions for manipulating the current address(/seal) of a capability(/set of seals): \texttt{cca} changes the current address(/seal) of a capability(/set of seals) by some offset, and \texttt{seta2b} sets the current address(/seal) of a capability(/set of seals) to the base address(/seal).
 The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
@@
 Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two rather than reducing the range of authority (\texttt{split}).
-The reverse is possible using \texttt{splice}.
-Sealables can be sealed using \texttt{cseal} and pairs of sealed capabilities can be unsealed by crossing security boundaries (\texttt{xjmp}, see below).
+\trgcm{} also has an instruction that does the inverse of splitting, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
+\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities by crossing security boundaries (\texttt{xjmp}, see below).
 Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).
@@
 For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In \stktokens{}, the base address of the stack is fixed as a compile-time constant.
-The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, as it was for the return token they gave to the callee.
+In \stktokens{}, the caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to the base address of the return token they gave to the callee.
 In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
@@
 % KJAA: It is unclear to me what 'at this point' refers to.
-% At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
-% However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
-% In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
-% To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
-% The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
-% The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
-% Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
-% At this point, the adversary has everything they need to return from either of the two callbacks.
-% Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
-% From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
-% They are both correct as the adversarial component has created two call stacks, one for each of the components.
-% In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
-% We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
-% A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
-% This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
-
-In \stktokens{} we ensure that at the start of execution, the stack memory is only referenced by a single, linear stack capability.
-Because of this, we can assert validity of the return token simply by checking its base address and splicing it with the caller's stack frame.
-There is no need to check linearity, because we know that only linear capabilities to this memory exist.
+At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
+However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
+In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
+To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
+The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
+The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
+Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
+At this point, the adversary has everything they need to return from either of the two callbacks.
+Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
+From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
+They are both correct as the adversarial component has created two call stacks, one for each of the components.
+In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
+We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
+A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
+This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
+
+Having a globally decided stack address also solves another challenge.
+Namely, the stack capability has to be linear in order for a caller to use it as a token, so a caller has to make sure that a capability is in fact linear when it is proposed as the stack capability by an untrusted source.
+This is, however, implicitly checked when the caller checks that the stack capability governs the global stack base address.
+The reason is that the stack capability needs to start execution as a linear capability which means that the capability that governs the global stack base address to begin with is linear.
+Due to the non-duplicability of linear capability and the fact that they cannot become non-linear during execution, the caller knows that if they have a capability that governs the global stack base address, then it must be linear.
 
@@
 
-By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack.
+By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from inactive stack frames on the call stack.
 Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack.
@@
 \todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
-Unfortunately due to space considerations, we cannot give a comprehensive description of it.
-Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensible description of a logical relation for a capability machine.
-Finally, for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.
+Unfortunately due to space considerations, we do not give a comprehensive description of it.
+Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
+Finally for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.
 
@@
 \label{subsec:worlds}
-Generally speaking, worlds are used in a logical relation to model the memory of a language.
-\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds should model this by being triples of sub-worlds.
-All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to some regions that depend on the kind of sub-world.
-where the regions differ for each of the sub-worlds.
-Generally, the regions consist of a world indexed relation that specifies the memories allowed by the region and a region type.
-The region type indicates how the region can be addressed which is important for linearity.
+Generally speaking, a world is used to model the memory of a language in a logical relation.
+\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
+All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
+Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
+The region types models linearity as they indicate how the memory can be addressed.
 We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
 The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
-The $\spatial$ region may seem odd as it not addressable, but all the relations in the logical relation are world indexed.
-If the register-file relation and the memory relation both have worlds with the same ownership, then they can both contain linear capabilities that address the same memory.
-If a spatial region is owned by the register-file relation, then it still has to be modelled in the memory which it can use a spatial region for.
-To capture this, we define a relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
-We also define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
-In a sense, linear capabilities are revocable, so the sub-worlds that allow spatial regions must also have a $\revoked$ region to model revocation.
+To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
+To this end, we have the relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
+%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
+The unaddressable $\spatial$ region makes sure that the memory it represents is modelled.
+Linear capabilities can be seen as revocable which is modelled by the $\revoked$ region.
 Finally, the $\pure$ region also contains a seal interpretation function.
 The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
-We split the regions in two categories ($\Wor$ given below):
+We split the regions in two categories:
 \begin{multline*}
@@
 \end{multline*}
-\begin{multline*}
-  \Regions = 
-    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup \\
-    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
+\begin{align*}
+  \Regions ={} &
+    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
+    & \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
     \{\revoked\}
-\end{multline*}
+\end{align*}
 and define the sub-worlds as follows:
@@
 \end{align*}
-The world for the memory can contain every kind of region because it is addressable by both normal and linear capabilities.
-The call stack and free stack should only be addressable by linear capabilities, so they only allow regions from $\Regions$.
+The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
+The call stack and free stack is only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
 Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.
 
-The worlds $\World$ are defined as $(\Worldh \times \Worlds \times \Worldfs)$ and they are recursive as they contain world indexed relations\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
-Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct them in the lemma below.
+Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$.
+They are recursive as they contain world indexed relations\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
+Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct them.
 \begin{lemma}
   \label{thm:recursive-domain-eq}
-  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ and there exists an isomorphism $\xi$ such that
+  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
   \[
@@
 \end{lemma}
-The future world relation $\future$ is fairly standard as it is extensional and requires regions to stay the same with the exception that $\spatial$ regions can become $\spatialo$ or $\revoked$ in a future world.
+The future world relation $\future$ is fairly standard.
+It is extensional and requires regions to stay the same in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.
 
@@
   \label{lem:compat-context-plug}
-  If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
-  $\npair[n']{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
+  If $\npair{\stpair{\context}{\context}}\in\lrcomp(W_1)$ and $\npair{\stpair{\comp}{\comp}} \in \lrcomp(W_2)$, then
+  $\npair[n']{(\plug{\context_S}{\comp_S},\plug{\context_T}{\comp_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
 \end{lemma}
@@
 
+\renewcommand{\comp}{C}
 \begin{figure}
@@
 
-    \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
       $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
-    \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
       $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
@@
 \label{sec:discussion}
-\begin{itemize}
-\item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
-\item Sharing stack references accross component boundaries is supported
-\item Other notions of well-bracketedness (specifically one would be to allow different stacks)
-\end{itemize}
-
+% \begin{itemize}
+% \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
+% \item Sharing stack references accross component boundaries is supported
+% \item Other notions of well-bracketedness (specifically one would be to allow different stacks)
+% \end{itemize}
+\subsection{Full abstraction}
+% - Full abstraction proofs difficult
+% - Need to compile to a machine with enforcement mechanisms - capability machine an option
+% - Full abstraction proofs modular, so other full abstraction proofs could target \srccm{} and thus have more abstractions to work with than if \trgcm{} was the target.
+
+\subsection{Sharing the stack}
+% ?
 \subsection{Practical applicability}
@@
 Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
-As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
+As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like $\tmove{}$, $\tstore{}$ and $\tload{}$.
 Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
-Our instructions are a bit clumsy from this point-of-view because, for example, \texttt{move} or \texttt{store} will zero the source register resp.
-memory location if the value being written is linear.
+Our instructions are a bit clumsy from this point-of-view because, for example, $\tmove{}$ or $\tstore{}$ will zero the source register resp. memory location if the value being written is linear.
 A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
@@
 
-Another problem is caused by the move semantics for \texttt{load} in the presence of multiple hardware threads.
+Another problem is caused by the move semantics for $\tload{}$ in the presence of multiple hardware threads.
 In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
-This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
+This means that a $\tload{}$ of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
+This is in principle not a problem except that a compare-and-swap is significantly slower than a regular $\tload{}$ (on the order of 10x slower or more).
 When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
 Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
-If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst-case...
+If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than $\tload$'s worst-case...
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten.

--- 2018-07-10 fully-abstract-comp/popl2019/paper.tex
@@
 The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
-In order for \stktokens{} to provide any guarantees for a component, we expect it to not shoot itself in the foot and of course perform the necessary stack check not captured by the call code (Figure~\ref{fig:call-code}).
+In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
 We more precisely expect four things of a reasonable component: (1) It checks the stack base address before performing a call.
@@
 \subsection{Full abstraction}
-All that is left before we state the full-abstraction theorem is to define how programs are lifted to executable configurations and contextual equivalence for each of the two capability machines we have defined.
+All that is left before we state the full-abstraction theorem is to define how components are combined with contexts and executed, so that we can define contextual equivalence.
 
 % How a program is lifted to an executable configuration
-Given a program as defined in Definition~\ref{def:program-and-context}, we want to be able to execute it, meaning we have to lift it to an executable configuration.
-When we lift a program to an executable configuration, we need to provide it a stack in order for \stktokens{} to work.
-The relation $(\comp_0, c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow \Phi$ lifts a component with a main to an executable configuration.
+Given a program $\var{comp}$, the judgement $\var{comp} \rightsquigarrow \Phi$ defines an initial execution configuration that can be executed.
 It works almost the same on \trgcm{} and \srccm{}.
-On both machines a stack that contains all zeroes is added.
-On \trgcm{} it is added as part of the memory along with the data and code memory.
-On \srccm{} it is added as the free stack in the configuration, and the initial stack is empty as no calls have been made.
+On both machines a stack containing all zeroes is added, as part of the regular memory on \trgcm{} and as the free stack on \srccm{}.
+On \srccm{}, the initial stack is empty as no calls have been made.
 The component needs access to the stack, so a stack pointer is added to the register file in $\rstk$.
@@
 The entry point of the program is specified by main, so the two capabilities are unsealed (they must have the same seal) and placed in the $\pcreg$ and $\rdata$ registers.
-The remainder of the registers contain zero.
+Other registers are set to zero.
 
@@
   \begin{gather*}
-    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2}
+    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2} \qedhere
   \end{gather*}
@@
 We split the regions in two categories ($\Wor$ given below):
-\begin{multline*}
-  \Regionh = 
+
+\begin{align*}
+\Regionh &= 
+\left\{\begin{multlined}
   \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
   (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
-\end{multline*}
-\begin{multline*}
-  \Regions = 
+\end{multlined}\right.\\
+  \Regions &= 
+\left\{\begin{multlined}
     \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup \\
@@
     \{\revoked\}
-\end{multline*}
+\end{multlined}\right.
+\end{align*}
 and define the sub-worlds as follows:
@@
     \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr),
-      ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W)
+      ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
   \]
@@
 Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
-If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst-case...
+If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
compilation issue

--- 2018-07-10 fully-abstract-comp/popl2019/paper.tex
@@
 The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability instructions simply inspect the properties of capabilities: type (\textt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
+The simplest of the capability instructions simply inspect the properties of capabilities: type (\texttt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
 The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Small corrections here and there

@@
 % xjmp
-Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
+Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them, and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
 A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
@@
 Opaque sealed capabilities encapsulate a security domain's local state and authority, and it only becomes accessible again when control is transferred to the security domain.
-Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one capability with the environment of another.
+Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one closure with the environment of another.
 % xjmp to unseal
@@
 In \stktokens{}, the base address of the stack is fixed as a compile-time constant.
-The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, as it was for the return token they gave to the callee.
+The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
 In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
@@
 
-In \stktokens{} we ensure that at the start of execution, the stack memory is only referenced by a single, linear stack capability.
+In \stktokens{}, we ensure that at the start of execution the stack memory is only referenced by a single linear stack capability.
 Because of this, we can assert validity of the return token simply by checking its base address and splicing it with the caller's stack frame.
-There is no need to check linearity, because we know that only linear capabilities to this memory exist.
+There is no need to check linearity because we know that only linear capabilities to this memory exist.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Pass over/rewrite of worlds and code reg subsections.

@@
 Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$.
-They are recursive as they contain world indexed relations\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
-Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct them.
-\begin{lemma}
-  \label{thm:recursive-domain-eq}
-  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
-  \[
-    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
-  \]
-  and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
-\end{lemma}
+They are recursive as they contain world indexed relations.
+Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct $\Wor$ which solves the circularity\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
+% \begin{lemma}
+%   \label{thm:recursive-domain-eq}
+%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
+%   \[
+%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
+%   \]
+%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
+% \end{lemma}
 The future world relation $\future$ is fairly standard.
-It is extensional and requires regions to stay the same in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.
+It is extensional and requires regions to stay the same type in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.
 
 \subsubsection{Code region}
-The code region is a central region to the logical relation used to model the code memory of a component.
-It is especially interesting because it specifies what the return and closure seals in a component can be used to seal.
+The code region is a central region to the logical relation that models the code memory of a component.
+It is especially interesting because its seal interpretation function specifies what the return and closure seals in a component can seal.
 The code region is defined as follows:
@@
 The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
-The memory relation $H^\mathrm{code,\square} \; (\dots)$ is very selective and only accepts the pair $(\mscode,\mscode)$ and only on the following conditions: The code memory $\mscode$ must be well-formed according to $\vdash$.
-Further, all the words contained in $\mscode$ must be non-linear and in the value relation, but as we will see later we have two value relations, a trusted and an untrusted, and the value relation used depends on whether $\mscode$ is trusted.
-For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have.
-The trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.
-If the addresses of $\mscode$ are in $\ta$, then the trusted value relation is used, and if $\mscode$ is disjoint with $\ta$, then the untrusted value relation is used.
-Further, if $\ta$ is untrusted, then $\sigrets$ has to be empty which means that $\mscode$ cannot contain return seals.
+The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself which corresponds with our component setup where code memory is read-only.
+$H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
+As we will see later, we have two value relations, a trusted and an untrusted, and the value relation used depends on whether the code region governs trusted or untrusted code.
+For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.
 
@@
 
-The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ in the code region dictates the return seals and closure seals can be used to seal.
-the return seals can only be used to seal return pointers, but only when they actually represent meaning full returns.
-for instance, a code return pointer should actually be associated with a trusted call, i.e.\ it should point to the address just after a call.
-the corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to do a couple of things to complete the call safely.
-further, the call they are associated with has to be the call that uses the seal in question.
-the code return pointer is not linear which means that it can be reused which is fine as it can only be used to return from the same call.
-on the other hand, the data return pointers are linear which is what prevents reuse of a return pair.
-multiple nested calls from the same trusted call will result in multiple stack frames that are each sealed with the same return seal.
-the seal interpretation function should allow a data return pointer for any of those pointers, so it simply requires that there exists a region $\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}$ (a region that only accepts $(\ms_S,\ms_T)$ and requires the contents to be in the untrusted value relation) in the part of the world that governs the call stack.
-The data return pointer should point to that stack frame.
-A region in the world for the call stack is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
-\todo[inline]{Consider writing something about how this relates to prove? We have to show that there is a stack frame, but we can rely on it when we receive a capability with the correct seal.}
+The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigclos$ can seal.
+The return seals can only be used to seal return pointers that corresponds to the return from an call in the code of the region.
+Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
+The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
+The code return pointer corresponds to a normal capability which means that it can be reused but always to return from the same call.
+Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
+Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function related data return pointers to linear capabilities.
+It also requires the data return pointer to correspond to one of the stack frames on the call stack.
+Which is expressed by requiring the existence of a spatially owned region in $\Worlds$ that governs memory corresponding to the return pointers range of authority.
+It must be spatially owned, so it is permitted to address it.
+A region in $\Worlds$ is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; (\ta,\stkb,\gsigrets,\gsigcloss) \; \sigma \; \hat{W}= \\
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \; \sigma \; \hat{W}= \\
   \begin{array}[t]{l}
@@
     \begin{array}{l}
-      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr'+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \mid 
-        \sigrets \subseteq \gsigrets \tand \\
-        \quad\dom(\mscode) \subseteq \ta \tand 
-        \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
-        \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
-        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
+      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\mathrm{xjmp\_off}+1)\right)} \mid \dots
+    %     \sigrets \subseteq \gsigrets \tand \\
+    %     \quad\dom(\mscode) \subseteq \ta \tand 
+    %     \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+    %     \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
+    %     \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
+    %     \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
     \end{array}
@@
     \begin{array}{l}
-      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid
-        \sigrets \subseteq \gsigrets \tand 
-        \dom(\mscode) \subseteq \ta \tand\\
-        \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
-        \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
-        \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
-        \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
+      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid \dots
+        % \sigrets \subseteq \gsigrets \tand 
+        % \dom(\mscode) \subseteq \ta \tand\\
+        % \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
+        % \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        % \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
+        % \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        % \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
     \end{array}
@@
 \end{multline*}
-We leave out the formal definition of $H^\mathrm{code,\square}_\sigma \; (\dots)$ for $\sigma \in \sigcloss$ and provide an informal description.
-What closure seals can be used for depends on whether the code is trusted.
-If the code is untrusted, i.e.\ $\dom(\mscode) \# \ta$, then the seal can be used for everything in the untrusted value relation can be sealed which correspond to everything the code has access to.
-On the other hand, if the code is trusted, then it depends on whether the sealable in question is executable.
-If it is executable, then it can be in the trusted value relation because we want to allow sealing of capabilities for the trusted code; but if it isn't executable, then the sealable has to be in the untrusted value relation.
+
+Unlike return seals, closure seals are allowed in both trusted and untrusted components, but we cannot allow them to use closure seals to seal the same things.
+If we allowed this, then a trusted component could seal a read-only capability for its code (i.e.\ access to return seals) with an untrusted closure seal and give it to the untrusted component which can unseal it with \texttt{xjmp}.
+The closure seals in untrusted code should therefore only be used to seal things from the untrusted value relation, i.e.\ things that are safe to transfer to an untrusted component.
+The trusted closure seals can by used to seal things from the untrusted value relation.
+This allows trusted seals to seal capabilities for the trusted code which is necessary in order to create a new closure.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Removed stack sharing from discussion.
More work on rewriting section 5.

@@
 \subsubsection{Expression relations and friends}
-The expression relation $\mathcal{E}$ in this logical relation contains pairs of capabilities that can safely be used for execution with respect to the give world.
-That is when capabilities are plugged into configurations with related register files  and related memory relative to the world, then the executions observably behave the same way.
-% the other expression relation
-With sealed capabilities in the mix, we will also need to be able to say something similar to the above but with a specific data capability in the register file.
-To this end, we have the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$ that relates two pairs of capabilities and requires that they are safe to use for the \texttt{xjmp} instruction.
+The logical relation has two expression relations.
+They both relate capabilities that can safely be used for execution in configurations that satisfy a given world.
+The first one $\mathcal{E}$ is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
+In our case, the only observable thing on the capability machines is termination.
+The observation relation $\mathcal{O}$ defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations.
+However, the only difference between the two approximations is in the observation approximation.}.
+\begin{align*}
+  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+  \forall i \leq n \ldotp
+    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
+  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+  \forall i \leq n \ldotp 
+    \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
+\end{align*}
+With sealed capabilities in the mix, we also need to be able to say that \texttt{xjmp}s are safe, but a sealed code capability only needs to be safe to execute when it has access to the related data capability.
+Further, \texttt{xjmp}s are used to invoke return pairs which on \srccm{} is somewhat involved.
+This is handled uniformly by the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$.
+It relates two pairs of capabilities by plugging them into related register-files and memories using the $\xjmpres{}$ function and requiring that the result is in the observation relation.
+\extend{Write that it is untrusted register files + intuition for this}
   % \begin{multline*}
@@
   % \end{multline*}
-\begin{multline*}
-  \lrexj(W) =
-  \\\left\{\begin{array}{l}
-      \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
-      \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
-      \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-    \end{array}
-    \right\}
-\end{multline*}
+% \begin{multline*}
+%   \lrexj(W) =
+%   \\\left\{\begin{array}{l}
+%       \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
+%       \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
+%       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+%       \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
+%       \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+%     \end{array}
+%     \right\}
+% \end{multline*}
 % register file relation
-The register-files relation $\lrr(\{\rdata\}) (W_\lrrs )$ relates register files when their registers, with register $\rdata$ and $\pcreg$ as the exception, contain words related in the untrusted value relation.
+
+Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
+The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, the they cannot depend on the same spatially owned region.
 % memory relation
-A memory on \trgcm{} $\ms_T$ is related to a heap, call stack and free stack on \srccm{} if the following is possible: $\ms_T$ can be split in three disjoint parts.
-One part relates to $\ms_S$ with respect to the heap world in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
-The second part should relate to the free stack $\ms_\stk$ also in the standard way, but with respect to the free stack world.
-The final part relates to the call stack $\stk$ which means that it can be further split into parts that each are related to one of the stack frames of $\stk$ with respect to the call stack world.
-The partition must be done in a way that preserves the order of the stack frames in $\stk$.
-% , so the top stack frame relates to a part of memory where all the other parts are above in the address space, the second stack frame relates to a part of memory where all the other parts are above in the address space with the exception of the topmost part, and so on.
-% Further, all stack frames should be in addresses above $\stkb$.
-The seal interpretation functions are in the pure regions, so the heap relation ($\lrheap$) also makes sure that only one seal interpretation function has an interpretation for each seal.
-\extend{Write about the particulars of the three memory satisfaction: heap relation makes sure only one interpretation for each seal. Free stack $\stkb$}
-The \texttt{xjmp} expression relation requires register files and memories related under different worlds, namely $W_\lrrs$ and $W_\lrm$, that are related with $\oplus$.
-This means that the two worlds have the same regions, but distribute the ownership of partial regions between them.
-By doing so, they make sure that for any part of memory only one of them can have a linear capability it.
-The two pairs of capabilities related by $\lrexj$ are plugged into their respective configurations using the $\xjmpres{}$ function, so it corresponds to how an $\texttt{xjmp}$ would insert them.
-% observation relation
-Finally, the observably accepted behaviour is defined by the observation relation $\lro$.
-\begin{align*}
-  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
-  \forall i \leq n \ldotp
-    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
-  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
-  \forall i \leq n \ldotp 
-    \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
-\end{align*}
-The only observable behavior of the capability machines is termination. 
-This means that the observation relations say that if one configuration terminates successfully, then the other one must terminate successfully as well.
-The fact that the observation relation is split into two parts means that we really have a logical approximation where one machine approximates the behavior of the other machine.
+The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
+This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that corresponds with the view \srccm{} provides on the memory.
+Each of the three partitions of $\ms_T$ must be related to the three parts of \srccm{} memory under the relevant sub-worlds of $W$.
+The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
+The free stack $\ms_\stk$ relates to the free stack partition of $\ms_T$ in the standard way, but with respect to the free stack world.
+%Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
+The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts relating each of the stack frames.
+Further, as each of the stack frames correspond to a slot in the return order, none of the stack frames should be empty.
+It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved in the actual memory.
+
+% %the topmost stack frame should have memory addresses smaller than (the stack grows downwards) the addresses all of the other stack frame, the next and so on.
+% A memory on \trgcm{} $\ms_T$ is related to a heap, call stack and free stack on \srccm{} if the following is possible: $\ms_T$ can be split in three disjoint parts.
+% One part relates to $\ms_S$ with respect to the heap world in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
+% The second part should relate to the free stack $\ms_\stk$ also in the standard way, but with respect to the free stack world.
+% The final part relates to the call stack $\stk$ which means that it can be further split into parts that each are related to one of the stack frames of $\stk$ with respect to the call stack world.
+% The partition must be done in a way that preserves the order of the stack frames in $\stk$.
+% % , so the top stack frame relates to a part of memory where all the other parts are above in the address space, the second stack frame relates to a part of memory where all the other parts are above in the address space with the exception of the topmost part, and so on.
+% % Further, all stack frames should be in addresses above $\stkb$.
+% The seal interpretation functions are in the pure regions, so the heap relation also makes sure that only one seal interpretation function has an interpretation for each seal.
+% \extend{Write about the particulars of the three memory satisfaction: heap relation makes sure only one interpretation for each seal. Free stack $\stkb$}
+% The \texttt{xjmp} expression relation requires register files and memories related under different worlds, namely $W_\lrrs$ and $W_\lrm$, that are related with $\oplus$.
+% This means that the two worlds have the same regions, but distribute the ownership of partial regions between them.
+% By doing so, they make sure that for any part of memory only one of them can have a linear capability it.
+% The two pairs of capabilities related by $\lrexj$ are plugged into their respective configurations using the $\xjmpres{}$ function, so it corresponds to how an $\texttt{xjmp}$ would insert them.
 
@@
 
-\begin{multline*}
-  \lrv[\gc=(\ta,\stkb,\gsigrets,\gsigcloss)](W) = \\
-  \begin{array}[t]{l}
-%   \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+\begin{align*}
+  \lrv(W) ={} & \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\ &
 %
     \left\{
-    \begin{array}{l}
-      \npair{\left(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr) \right)} \; | \;
-        \perm \not\in \{\rx,\rwx\} \tand \\
-        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in
-        % \lrp(W) \wedge \\
-        \quad\perm \in \{\ro,\rw\} \Rightarrow \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \tand \\
-        \quad\perm = \rw  \Rightarrow \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
-    \end{array}
-    \right\} \cup \\
+%    \begin{array}{l}
+      \npair{\left(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr) \right)} \mid\dots
+        % \perm \not\in \{\rx,\rwx\} \tand \\
+        % % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in
+        % % \lrp(W) \wedge \\
+        % \quad\perm \in \{\ro,\rw\} \Rightarrow \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \tand \\
+        % \quad\perm = \rw  \Rightarrow \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
+%    \end{array}
+    \right\} \cup \\ &
 %
     \left\{
-    \begin{array}{l}
-    \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)}\; | \; [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
-      \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
-    \end{array}
-    \right\} \cup \\
+%    \begin{array}{l}
+      \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} \mid 
+      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss) \wedge \dots 
+      % \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
+      % \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+%    \end{array}
+    \right\} \cup \\ &
         \left\{
-    \begin{array}{l}
-      \npair{\left(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T} \right)} \; | \; \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
-      \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
-      \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \tand \dots
-      % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-    \end{array}
-    \right\}\cup \dots
+%    \begin{array}{l}
+      \npair{\left(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T} \right)} \mid \dots
+      % \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
+      % \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      % \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      % \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
+      % \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
+      % \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      % \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \tand \dots
+      % % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
+      % % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      % % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+%    \end{array}
+    \right\}\cup\\ &
 %
-    % \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+     \left\{ \npair{\left(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr)\right)} \mid \dots
     % \begin{array}{l}
@@
     % \end{array}
-    % \right\}
-  \end{array}
-\end{multline*}
+     \right\}
+%  \end{array}
+\end{align*}
 A stack pointers is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
@@
 The trusted value relation includes everything that the untrusted value relation contains and a few extra things.
-\begin{multline*}
-  \lrvtrusted[\square,\gc=(\ta,\stkb,\gsigrets,\gsigcloss)](W) = \lrv(W)\cup \\
-  \begin{array}[t]{l}
-    \left\{ \begin{array}{l} \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} 
-    \; | \; \exists r \in \dom(\pwheap) \ldotp \\
-              \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\mscode,\gc} \tand \dom(\mscode) \subseteq \ta \\
-              \quad \tand [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \tand \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss
-    \end{array}
+
+\begin{align*}
+  \lrvtrusted(W) ={} & \lrv(W)\cup \\
+%  \begin{array}[t]{l}
+    &\left\{
+    %\begin{array}{l}
+      \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} \mid
+      [\sigma_\baddr,\sigma_\eaddr] \subseteq(\sigrets \cup \sigcloss) \wedge \dots 
+    %           \exists r \in \dom(\pwheap) \ldotp \\
+    %           \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\mscode,\gc} \tand \dom(\mscode) \subseteq \ta \\
+    %           \quad \tand [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \tand \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss
+    % \end{array}
     \right\} \cup \\
-    \left\{
-    \begin{array}{l}
-      \npair{\left(\src{((\perm,\normal),\baddr,\eaddr,\aaddr)},((\perm,\normal),\baddr,\eaddr,\aaddr) \right)} \; |  \\
-      \quad \perm \sqsubseteq \rx \tand 
-       [\baddr,\eaddr] \subseteq \ta \tand 
-       \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
-    \end{array}
+    & \left\{
+%    \begin{array}{l}
+      \npair{\left(\src{((\perm,\normal),\baddr,\eaddr,\aaddr)},((\perm,\normal),\baddr,\eaddr,\aaddr) \right)} \mid \perm \le \rx \wedge \dots
+    %   \quad \perm \sqsubseteq \rx \tand 
+    %    [\baddr,\eaddr] \subseteq \ta \tand 
+    %    \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
+    % \end{array}
     \right\}
-  \end{array}
-\end{multline*}
+%  \end{array}
+\end{align*}
 A side from the sets of seals in the untrusted value relation, sets of return and closure seals are in the trusted value relation when they belong to some trusted code.
@@
 Generally for executable code capabilities, the $\xReadCond{}$ should hold which requires there to be a code region that governs the piece of code memory that the capability has authority over.
-\[
-  \xReadCond{W} = \left\{ \npair{A} \middle|
-    \begin{array}{l}
-      \exists r \in \addressable{\normal, \pwheap} \ldotp \\
-       \pwheap(r) \nequal \codereg{\_,\_,\mscode,\gc}\tand
-       \dom(\mscode) \supseteq A 
-    \end{array}
-  \right\}
-\]
+% \[
+%   \xReadCond{W} = \left\{ \npair{A} \middle|
+%     \begin{array}{l}
+%       \exists r \in \addressable{\normal, \pwheap} \ldotp \\
+%        \pwheap(r) \nequal \codereg{\_,\_,\mscode,\gc}\tand
+%        \dom(\mscode) \supseteq A 
+%     \end{array}
+%   \right\}
+% \]
+
+\todo[inline]{Write some intuition about monotonicity, owned spatial regions and revoked regions}
 
@@
 \end{lemma}
-It roughly says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
+The FTLR roughly says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
 This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ ensures, and for trusted code the semantic reasonability condition must also be satisfied by the code.
@@
 % - Full abstraction proofs difficult
+
 % - Need to compile to a machine with enforcement mechanisms - capability machine an option
@@
 
-\subsection{Sharing the stack}
+%\subsection{Sharing the stack}
 % ?

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
value relation subsubsection rewritten.

@@
 The purpose of the value relations is two fold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
-As previously explained, we try to ensure LSE and WBCF for some trusted code of which we require certain things like treating return seals in a reasonable way.
-This means that it is safe to give return seals to trusted code.
-At the same time, we have some code that we do not trust, and given a return seal it may try to abuse it to break LSE and WBCF for the trusted code.
+As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonable, e.g.\ by treating return seals in a reasonable way.
+This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.
 This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and untrusted value relation.
-
+The two value relations are sketched in Figure~\ref{fig:value-relation}.
+
+The untrusted value relation relates words that can safely be given to untrusted code.
+Data is always safe, and a piece of data relates to itself.
+A stack pointers is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
+The stack pointer should only be able to produce safe words, so a stack pointer with read permission can only produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a stack pointer with write permission can at least write safe words.
+Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack part of the world.
+Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read-write permission.
+The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
+An executable memory capability should always point to code which means that there should be a code region governing the memory it points to.
+The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seal is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
+% The write condition is called $\stackWriteCond{}$
+% \begin{multline*}
+%   \stackWriteCond{W} =\\
+%   \left\{
+%     \begin{array}{l}\npair{A} \mid 
+%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
+%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
+%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
+%     \end{array}
+%   \right\}
+% \end{multline*}
+When a piece of untrusted code is called, it is passed a return pair consisting of two sealed capabilities.
+Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
+The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\vsc_S,\vsc_T)$ must be in the seal interpretation function for $\sigma$.
+Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.
+
+\begin{figure}[htb]
+  \centering
   \begin{align*}
@@
     % \end{array}
-     \right\}
+     \right\} \\
 %  \end{array}
-\end{align*}
-A stack pointers is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
-In addition depending on the permission, a read and write condition needs to be satisfied.
-Intuitively, the read condition guarantees that the capability can only be used to read safe values, i.e.\ values in $\mathcal{V}_\untrusted$.
-Similarly the write condition says that the capabilities can at least be used to write the safe values.
-These two properties are expressed by picking out the regions from the world that govern the memory the stack pointer gives access to.
-The free stack is modelled by the free stack world, so the regions must come from that part of the world.
-Further, stack pointers are linear, so the regions they depend on must be addressable which means that they must be owned by that free stack world.
-% The write condition is called $\stackWriteCond{}$
-% \begin{multline*}
-%   \stackWriteCond{W} =\\
-%   \left\{
-%     \begin{array}{l}\npair{A} \mid 
-%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
-%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
-%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
-%     \end{array}
-%   \right\}
-% \end{multline*}
-
-The untrusted value relation also relates sets of seals that are syntactically identical and that are disjoint from the return and closure seals of interest ($\gsigrets$ and $\gsigcloss$). 
-Further, the seals should only be used to seal words from the untrusted value relation, so the seal interpretation function for each of the seals must be the untrusted value relation.
-
-The untrusted value relation also needs to include sealed capabilities in order to allow return pairs and closures to be given to untrusted code.
-The sealed capabilities should of course be sealed with the same seal, and the seal interpretation function for that seal should be equal to the one of the code region.
-Aside from this, the pair of sealables sealed by the seal should satisfy two things: (1) they should be in the seal interpretation function for that seal and (2) in any possible future\footnote{The $\purePart{}$ function strips away all ownership from a world.
-A non-linear capability should not depend on linear things, so we take any future of $\purePart{W}$.}, it should be safe to jump to the sealable pair when they are paired with sealables permitted by the seal interpretation function for that seal, so the quadruple must be in the $\mathcal{E}_\mathrm{xjmp}$ relation.
-
-We have left out the part of the untrusted value relation that relates memory capabilities and integers (they are related when they are equal).
-The conditions for related memory capabilities are almost the same as for stack pointers.
-The main difference is that the conditions depend on the heap part of the world rather than the free stack.
-Further, memory capabilities can be executable which needs extra care to ensure security which we will see in the trusted value relation.
-
-The trusted value relation includes everything that the untrusted value relation contains and a few extra things.
-
-\begin{align*}
   \lrvtrusted(W) ={} & \lrv(W)\cup \\
@@
 \end{align*}
-A side from the sets of seals in the untrusted value relation, sets of return and closure seals are in the trusted value relation when they belong to some trusted code.
-This means that in the heap part of the world there must be a code region that has the seals in question, and the code should be in $\ta$.
-
-The trusted value relation also contains executable capabilities for the trusted code.
-Generally for executable code capabilities, the $\xReadCond{}$ should hold which requires there to be a code region that governs the piece of code memory that the capability has authority over.
+\caption{Sketches of the trusted and untrusted value relation.}
+\label{fig:value-relation}
+\end{figure}
+
+The trusted value relation includes everything that the untrusted value relation contains and a few extra things.
+The trusted code uses the return seals and the closure seals in its program, and it does so in a responsible way which is why sets of seals with trusted return seals and closure seals are in the trusted value relation.
+In order for trusted code to even execute, it needs to use an executable memory capability for the trusted code, so such capabilities are in the untrusted value relation.\footnote{Note that code capabilities for trusted code is not allowed in the untrusted value relation as it could be used to read return seals that are not in the untrusted value relation.}
+
 % \[
@@
 
-\todo[inline]{Write some intuition about monotonicity, owned spatial regions and revoked regions}
-
-With the logical relation defined, we are ready to state the fundamental theorem of logical relations (FTLR).
+\extend{Write some intuition about monotonicity, owned spatial regions and revoked regions}
+
+\subsubsection{Fundamental theorem}
+With the logical relation defined, we state the fundamental theorem of logical relations (FTLR).
 \begin{lemma}[FTLR]

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
FTLR subsubsection updated.

@@
   \[
-    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr),
-      ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W)
+    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr), ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W)
   \]
 \end{lemma}
-The FTLR roughly says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
-This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ ensures, and for trusted code the semantic reasonability condition must also be satisfied by the code.
-The proof of the FTLR is done by a nested induction that categorises and considers each possible step the two executions can take.
-While the lemma is not used directly in the proof sketch we present in Section~\ref{subsec:proof-sketch}, it is detrimental to the proof of the lemmas we use in it.
+Roughly speaking, the FTLR says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
+This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ entails, and for trusted code the semantic reasonability condition must also be satisfied.
+The proof of the FTLR is done by a nested induction where each instruction that could be executed is considered.
+While the lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, it is detrimental to the proofs of the lemmas in that section.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Formatting of lemma.
Small changes to the last subsubsection of section 5.

@@
 The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward.
-The component relation relates components to them self which makes sense as well-formed components should contain no \srccm{} specific constructs.
+The component relation relates components to themselves which makes sense as well-formed components should contain no \srccm{} specific constructs.
 We prove a number of lemmas about the relations.
@@
   \label{lem:ftlr-comps}
-  If
-  \begin{itemize}
-  \item $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$
-  \item One of the following holds:
-    \begin{itemize}
-    \item $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component
-    \item $\dom(\comp.\mscode) \mathrel{\#} \ta$
-    \end{itemize}
-  \end{itemize}
-  Then there exists a $W$ such that
+  If $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$ and either
+    $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component; or
+    $\dom(\comp.\mscode) \mathrel{\#} \ta$,
+  then there exists a $W$ such that
   $\npair{(\comp,\comp)} \in \lrcomp(W)$
@@
   The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.
-  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that $\comp_1$ and $\comp_2$ are well-formed and reasonable.
-  Further, by $\tconeq$ the context $\context$ must be well-formed.
-  We prove arrow (1) in the figure by using the mentioned assumption about $\comp_1$ and $\context$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
-  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$, so by the assumption that $\comp_1$ and $\comp_2$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ which proves arrow (2) in the figure.
-  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use that $\comp_2$ is well-formed and reasonable and that $\context$ is well-formed.
+  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that the trusted components $\comp_1$ and $\comp_2$ are well-formed and reasonable.
+  Further, by $\tconeq$ we may assume the context $\context$ is well-formed.
+  We prove arrow (1) in the figure by using the mentioned assumptions about $\comp_1$ and $\context$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
+  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$, so by the assumption that $\comp_1$ and $\comp_2$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$, i.e.\ arrow (2) in the figure.
+  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use the assumption that $\comp_2$ is well-formed and reasonable and that $\context$ is well-formed.
 \end{proof}

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Full abstraction discussion (1st draft)

@@
 % - Full abstraction proofs difficult
+In the grand scheme of proving full abstraction from a high-level language to assembly, we believe \stktokens{} can play a role.
+The target of a fully-abstract compiler must have enforcement mechanisms that can enforce the properties of the high-level language.
+As demonstrated in this paper, a capability machine with linear capabilities can enforce two such properties namely LSE and WBCF.
+Further due to the modularity of full-abstraction proofs, fully abstract compilers from languages with LSE and WBCF can target \srccm{} directly and rely on the full-abstraction proof this paper presents.
+However in order for a fully abstract compiler utilise our full-abstraction proof, it must satisfy the conditions of the proof, namely the compiled code must be well-formed and well-behaved.
+From a high-level perspective this condition actually says that the compiled code does not need to protect itself from itself which it would have to without the condition.
+
+When creating fully-abstract compilers between low-level machines, it is a big challenge to work with the exposed addresses.
+In particular, if the compilation changes the code size of a block of code, then it may be observable and prevent full-abstraction from being proven.
+We would argue that when a compilation reaches a phase where addresses are exposed, then the compilation should no longer change the code.
+This does, however, pose the challenge that there might quite a few abstractions in difference between the language where addresses are hidden to a machine where they are not.
+We propose that this challenge is solved by implementing these abstractions in a number of overlay semantics.
+By implementing them one by one, one can deal with one abstraction at a time reducing the complexity of each necessary full-abstraction proof.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Contribution list.

@@
 
-\todo[inline]{Add itemized contributions (see comments). Outline?}
-% The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure local state encapsulation and well-bracketed control flow, and (2) a novel proof technique that we call fully-abstract overlay semantics.
-
+%\todo[inline]{Add itemized contributions (see comments). Outline?}
+%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
+In this paper, we make the following contributions:
 % Presents the first formal model of a capability machine with linear capabilities (Section 2)
@@
 % Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
-
-% \begin{itemize}
-% \item 
-% \end{itemize}
+\begin{itemize}
+\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
+\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
+\item We present and use a novel proof technique called fully-abstract overlay semantics which we use to prove the properties of \stktokens{}. This includes:
+  \begin{itemize}
+  \item defining \srccm{} an overlay semantics for \trgcm{} (Section~\ref{sec:form-secur-with}) and
+  \item proving a full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
+  \item using and defining cross language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
+\end{itemize}
+\end{itemize}
+The technical report~\citep{technical_report} contains additional details and proofs that we left out.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten references to TR.

@@
 
-Some simplifications has been made in this presentation of \trgcm{} in this section. We refer to the accompanying technical report\citep{technical_report} for all details.
+Some simplifications has been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
 \begin{figure}[htb]
@@
 The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the syntactic requirements necessary to be able to rely on system guarantees.
-For space reasons, we just describe the main points of the judgement and refer to the technical report~\citep{technical_report} for the formal definition.
+We elide the details of the judgement here and describe the main points.
 For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
@@
 By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
-The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.
+We elide the details here and refer to~\citet{technical_report}
+%The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Introduction to Section 5.

@@
 \label{sec:fa-proof}
-\todo[inline]{Write small introduction.
+%\todo[inline]{Write small introduction.
 Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
@@
 % \end{itemize}
+The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
+It relates the observable behavior of \srccm{} to the observable behavior of \trgcm{}.
+In the full-abstraction proof, it allows us to transfer termination assumptions between the two machines which enables the use of a contextual equivalence assumption.
+We first give an informal presentation of the logical relation before we return to a sketch of the full-abstraction proof.
 
@@
 \label{subsec:logical-relation}
-The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
-\todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
+In this section, we present the logical relation.
 Unfortunately due to space considerations, we do not give a comprehensive description of it.

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex

@@
 We split the regions in two categories:
-\begin{multline*}
-  \Regionh = 
+\begin{align*}
+  \Regionh &= 
+             \left\{\begin{multlined}
                \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
                (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
-\end{multline*}
-\begin{align*}
-  \Regions ={} &
+             \end{multlined}\right.\\
+  \Regions &={}
+             \left\{\begin{multlined}
                \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
-    & \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
+               \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
                \{\revoked\}
+             \end{multlined}\right.
 \end{align*}
@@
   \[
-    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr), ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W)
+    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr), ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
   \]

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Tune discussion of full abstraction a bit.

@@
 % - Full abstraction proofs difficult
-In the grand scheme of proving full abstraction from a high-level language to assembly, we believe \stktokens{} can play a role.
-The target of a fully-abstract compiler must have enforcement mechanisms that can enforce the properties of the high-level language.
-As demonstrated in this paper, a capability machine with linear capabilities can enforce two such properties namely LSE and WBCF.
-Further due to the modularity of full-abstraction proofs, fully abstract compilers from languages with LSE and WBCF can target \srccm{} directly and rely on the full-abstraction proof this paper presents.
-However in order for a fully abstract compiler utilise our full-abstraction proof, it must satisfy the conditions of the proof, namely the compiled code must be well-formed and well-behaved.
-From a high-level perspective this condition actually says that the compiled code does not need to protect itself from itself which it would have to without the condition.
-
-When creating fully-abstract compilers between low-level machines, it is a big challenge to work with the exposed addresses.
-In particular, if the compilation changes the code size of a block of code, then it may be observable and prevent full-abstraction from being proven.
-We would argue that when a compilation reaches a phase where addresses are exposed, then the compilation should no longer change the code.
-This does, however, pose the challenge that there might quite a few abstractions in difference between the language where addresses are hidden to a machine where they are not.
-We propose that this challenge is solved by implementing these abstractions in a number of overlay semantics.
-By implementing them one by one, one can deal with one abstraction at a time reducing the complexity of each necessary full-abstraction proof.
+Our formulation of WBCF and LSE using a fully abstract overlay semantics has an advantage with respect to others that we haven't discussed yet.
+Imagine that you are implementing a fully abstract compiler for a high-level language, i.e.\ a secure compiler that enforces high-level abstractions when interacting with untrusted target-language components.
+Such a compiler would need to perform many things and enforce other high-level properties than just WBCF and LSE.
+
+If such a compiler uses the \stktokens{} calling convention, then the security proof should not have to reprove security of \stktokens{}.
+Ideally, it should just combine security proofs for the compiler's other functionality with our results about \stktokens{}.
+We want to point out that our formulation enables such reuse.
+Specifically, the compiler could be factored into a part that targets \srccm{}, followed by our embedding into \trgcm{}.
+If the authors of the secure compiler can prove full abstraction of the first part (relying on WBCF and LSE in \srccm{}) and they can also prove that this first part generates well-formed and reasonable components, then full abstraction of the whole compiler follows by our result and transitivity of fully abstract compilation.
+Perhaps other reusable components of secure compilers could be formulated similarly using some form of fully abstract overlay semantics, to obtain similar reusability of their security proofs.
+
+% When creating fully-abstract compilers between low-level machines, it is a big challenge to work with the exposed addresses.
+% In particular, if the compilation changes the code size of a block of code, then it may be observable and prevent full-abstraction from being proven.
+% We would argue that when a compilation reaches a phase where addresses are exposed, then the compilation should no longer change the code.
+% This does, however, pose the challenge that there might quite a few abstractions in difference between the language where addresses are hidden to a machine where they are not.
+% We propose that this challenge is solved by implementing these abstractions in a number of overlay semantics.
+% By implementing them one by one, one can deal with one abstraction at a time reducing the complexity of each necessary full-abstraction proof.
 

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
Shorten a bit.

@@
 The trusted value relation includes everything that the untrusted value relation contains and a few extra things.
-The trusted code uses the return seals and the closure seals in its program, and it does so in a responsible way which is why sets of seals with trusted return seals and closure seals are in the trusted value relation.
+Trusted code uses the return seals and the closure seals (in a responsible way), and therefore sets of seals with trusted return seals and closure seals are in the trusted value relation.
 In order for trusted code to even execute, it needs to use an executable memory capability for the trusted code, so such capabilities are in the untrusted value relation.\footnote{Note that code capabilities for trusted code is not allowed in the untrusted value relation as it could be used to read return seals that are not in the untrusted value relation.}
@@
 The proof of the FTLR is done by a nested induction where each instruction that could be executed is considered.
-While the lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, it is detrimental to the proofs of the lemmas in that section.
+The lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, but it is crucial to the proofs of the lemmas in that section.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Removed resolved comments.

@@
 % The other attack.
-% KJAA: It is unclear to me what 'at this point' refers to.
 % At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
@@
 \item \label{item:non-empty-sf} Make sure that local stack frames are non-empty
-% KJAA: There's a comma (or something!) missing below. 
-\item \label{item:return-data} Create token and data return capability on call, split the stack capability in two to get a stack capability for your local stack frame, and a stack capability for the unused part of the stack.
+\item \label{item:return-data} Create token and data return capability on call: split the stack capability in two to get a stack capability for your local stack frame and a stack capability for the unused part of the stack.
 The stack capability for your local stack frame is sealed and used for the data part of the return pair.
-  % KJAA: Same with the last two items
-\item \label{item:return-code} Create code return capability on call Seal the old program pointer.
-\item Reasonable use of seals return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
+\item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
+\item Reasonable use of seals: Return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
 \end{enumerate}

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Added new first line to introduction.

@@
 \label{sec:introduction}
-\todo[inline]{Add a sentence or two about secure compilation to ease the start.}
+Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Spellcheck

@@
   The calling convention is based on linear capabilities, a type of capabilities that are prevented from being duplicated by the hardware.
-  In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, using what we call a fully abstract overlay semantics.
+  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, using what we call a fully abstract overlay semantics.
 \end{abstract}
@@
 Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To accomodate this, we define a second operational semantics for our target language, with a native well-bracketed stack and primitive ways to do calls and returns.
+To accommodate this, we define a second operational semantics for our target language, with a native well-bracketed stack and primitive ways to do calls and returns.
 This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
@@
 \]
-  \caption{The syntax of our capability machine with seals and linear capbilities.}
+  \caption{The syntax of our capability machine with seals and linear capabilities.}
   \label{fig:target-syntax}
@@
 The fact that \stktokens{} enforces LSE and WBCF is formulated as the theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
-The theorem states that this function consitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
+The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
 Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than the corresponding components in the well-behaved overlay semantics.
@@
 Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
-The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modeled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
+The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
 This is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
@@
 The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation at play.
-They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared accross component boundaries.
+They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared across component boundaries.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Grammar and typos

@@
   We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
-  The calling convention is based on linear capabilities, a type of capabilities that are prevented from being duplicated by the hardware.
-  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, using what we call a fully abstract overlay semantics.
+  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
+  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
 \end{abstract}
@@
 \label{sec:introduction}
+\todo[inline]{TODOS ENABLED!!!}
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
@@
 
-Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
+Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves, or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
 At the assembly level, this property does not hold immediately. 
@@
 However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
-To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
+To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
 The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
@@
 % KJAA: This 'sealed' and 'common seal' business is unclear to me.
-On a recent capability machine called CHERI, for example, they take the form of pairs of capabilities that represent the code and data parts of the closure.
+For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
 Each of the two capabilities are sealed with a common seal which make them opaque.
@@
 This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it scales poorly to large amounts of distrusting components, because of the need to reserve separate stack space for all components.
-Also, in the presence of higher-order values (e.g.
+First, it scales poorly to large amounts of distrusting components because of the need to reserve separate stack space for all components.
+Also in the presence of higher-order values (e.g.
 function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
@@
 A more scalable approach retains a single stack that is shared between components.
-Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component, and to revoke them after it returns.
+Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component and to revoke them after it returns.
 While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
-For example, CHERI offers a form of {\itshape local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
+For example, CHERI offers a form of {\itshape local} capabilities that, briefly, can only be stored in registers or on the stack but not in other parts of memory.
 \citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
@@
 % KJAA: Perhaps make it clear by who and when/where?
-These capabilities are non-duplicable, enforced by the hardware.
+These capabilities are non-duplicable enforced by the hardware.
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings, particularly without the need for clearing large blocks of memory.
-
-A second contribution of this work is the way in which we formulate these two properties, using a technique we call {\itshape fully abstract overlay semantics}.
+The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings particularly without the need for clearing large blocks of memory.
+
+A second contribution of this work is the way in which we formulate these two properties using a technique we call {\itshape fully abstract overlay semantics}.
 Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
 Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To accommodate this, we define a second operational semantics for our target language, with a native well-bracketed stack and primitive ways to do calls and returns.
+To accommodate this, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
 This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
@@
 \item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present and use a novel proof technique called fully-abstract overlay semantics which we use to prove the properties of \stktokens{}. This includes:
+\item We present and use a novel proof technique called \textit{fully-abstract overlay semantics} which we use to prove the properties of \stktokens{}. This includes:
   \begin{itemize}
   \item defining \srccm{} an overlay semantics for \trgcm{} (Section~\ref{sec:form-secur-with}) and
-  \item proving a full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
-  \item using and defining cross language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
+  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
+  \item using and defining a cross language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
 \end{itemize}
 \end{itemize}
-The technical report~\citep{technical_report} contains additional details and proofs that we left out.
+This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Moved figures around.

@@
 \label{sec:cap-mach-w-seal-and-lin}
+% Brief intro to section and this cap machine:
+% Cap machine inspired by CHERI, but with linear capabilities
+\begin{wrapfigure}{r}{0.28\linewidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (rwx) {$\rwx$};
+    \node[main node] (rx) [below right of=rwx] {$\rx$};
+
+    \node[main node] (rw) [below left of=rwx,] {$\rw$};
+    \node[main node] (r) [below right of=rw] {$\readonly$};
+    \node[main node] (0) [below of=r] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+
+    (rw) edge (r)
+    (r) edge (0)
+
+    (rwx) edge (rx)
+
+    (rw) edge (rwx)
+    (r) edge (rx);
+  \end{tikzpicture}
+
+  \caption{Permission hierarchy}
+  \label{fig:perm-hier}
+\end{wrapfigure}
+In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
+For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
+% \lau{Do we have a good reason to assume an infinite address space for this work?
+%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
+
+%%% Capabilities
+The concept of a capability is the cornerstone of any capability machine.
+In its simplest form, a capability is a permission and a range of authority.
+The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
+% Permissions and Range of authority
+The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}).
+Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
+The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
+% Pointer and linearity
+In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
+The linearity $\lin$ is either $\normal$ for traditional capabilities $\linear$ for linear ones.
+% What is linear capabilities, split/splice
+A linear capability is a capability that cannot be duplicated.
+This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source register or memory location is cleared.
+The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
+
+%%% Seals
+% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
+
+Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
+%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
+Syntactically, sealed capabilities add two types of capabilities.
+% seals
+A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
+% sealed caps
+A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
+Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
 \begin{figure}[tb]
@@
 \end{figure}
-% Brief intro to section and this cap machine:
-% Cap machine inspired by CHERI, but with linear capabilities
-In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
-For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
-% \lau{Do we have a good reason to assume an infinite address space for this work?
-%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
-
-%%% Capabilities
-The concept of a capability is the cornerstone of any capability machine.
-In its simplest form, a capability is a permission and a range of authority.
-The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
-% Permissions and Range of authority
-The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}).
-Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
-The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
-% Pointer and linearity
-In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
-The linearity $\lin$ is either $\normal$ for traditional capabilities $\linear$ for linear ones.
-% What is linear capabilities, split/splice
-A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source register or memory location is cleared.
-The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
-
-\begin{wrapfigure}{r}{0.28\linewidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (rwx) {$\rwx$};
-    \node[main node] (rx) [below right of=rwx] {$\rx$};
-
-    \node[main node] (rw) [below left of=rwx,] {$\rw$};
-    \node[main node] (r) [below right of=rw] {$\readonly$};
-    \node[main node] (0) [below of=r] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-
-    (rw) edge (r)
-    (r) edge (0)
-
-    (rwx) edge (rx)
-
-    (rw) edge (rwx)
-    (r) edge (rx);
-  \end{tikzpicture}
-
-  \caption{Permission hierarchy}
-  \label{fig:perm-hier}
-\end{wrapfigure}
-
-%%% Seals
-% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
-
-Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
-%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
-Syntactically, sealed capabilities add two types of capabilities.
-% seals
-A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
-% sealed caps
-A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
-Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Addressed Lars comment in introduction.

@@
   e.g., by contrasting it to our earlier work where the WBCF and LSE properties were only proved via specific examples.}
-This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
+Compared to~\citet{skorstengaard_reasoning_2017} that show LSE and WBCF in a handful of examples,
+this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
 Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Grammar+shorten.

@@
 
-Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves, or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
+Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
 At the assembly level, this property does not hold immediately. 
@@
 The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
-This is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
-For simplicity, we have omitted features of CHERI that were not needed for modelling \stktokens{} (e.g.\ local capabilities, virtual memory).
-
-Plenty of other papers enforce well-bracketed control flow at a low level, but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+This is a recent, relatively mature capability machine which combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
+For simplicity, we have omitted features of CHERI that were not needed for \stktokens{} (e.g.\ local capabilities, virtual memory).
+
+Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
 This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
@@
 Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
-However, its attacker model is different and its security goals are weaker.
-They assume an attacker that is not able to execute code, but can overwrite arbitrary data at any time during execution (to model buffer overflows).
+However, its attacker model is different, and its security goals are weaker.
+They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
 In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness, saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
+Also, it only enforces a weak form of control flow correctness saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
 Such a property ignores temporal properties and seems hard to use for reasoning.
@@
 More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016,juglaret_beyond_2016}.
-Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
-Both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
+Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments), and we do not use per-component stacks or a trusted stack manager but a single shared stack and a decentralized calling convention based on linear capabilities.
+Both prove a secure compilation result from a high-level language which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
 
@@
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler, where these scenarios are not relevant.
+Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Shortening.

@@
 
-A more scalable approach retains a single stack that is shared between components.
+A more scalable approach retains a single stack shared between components.
 Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component and to revoke them after it returns.
 While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
-For example, CHERI offers a form of {\itshape local} capabilities that, briefly, can only be stored in registers or on the stack but not in other parts of memory.
+For example, CHERI offers a form of {\itshape local} capabilities that can only be stored in registers or on the stack but not in other parts of memory.
 \citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
@@
 % Cap machine inspired by CHERI, but with linear capabilities
-\begin{wrapfigure}{r}{0.28\linewidth}
+\begin{wrapfigure}{r}{0.26\linewidth}
   \centering
@@
 % Permissions and Range of authority
-The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}).
+The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
 Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
@@
 In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
-The linearity $\lin$ is either $\normal$ for traditional capabilities $\linear$ for linear ones.
+The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
 % What is linear capabilities, split/splice
 A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source register or memory location is cleared.
+This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
 The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
@@
 Some simplifications has been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
-\begin{figure}[htb]
+\begin{figure}[tb]
   \centering
@@
 In our case, the only observable thing on the capability machines is termination.
-The observation relation $\mathcal{O}$ defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations.
+The observation relation $\mathcal{O}$ (Figure~\ref{fig:obs-rel}) defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations.
 However, the only difference between the two approximations is in the observation approximation.}.
+\begin{figure}[htb]
+  \centering
   \begin{align*}
@@
 \end{align*}
+\caption{The observation approximations.}
+\label{fig:obs-rel}
+\end{figure}
 With sealed capabilities in the mix, we also need to be able to say that \texttt{xjmp}s are safe, but a sealed code capability only needs to be safe to execute when it has access to the related data capability.
@@
 
-Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+Plenty of other papers enforce WBCF at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
 This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Remove resolved todos

@@
 
-\todo[inline]{In the below, Lars would like to be more precise about the ``in a general way'' below.
-  e.g., by contrasting it to our earlier work where the WBCF and LSE properties were only proved via specific examples.}
-Compared to~\citet{skorstengaard_reasoning_2017} that show LSE and WBCF in a handful of examples,
+Compared to~\citet{skorstengaard_reasoning_2017} that prove LSE and WBCF for a handful of examples,
 this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
@@
 
-%\todo[inline]{Add itemized contributions (see comments). Outline?}
 %The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
@@
 Components and linking are defined in Figure~\ref{fig:target-component-and-linking}.
-\todo[inline]{Add better separation between component and linking in figure.}
 
@@
 In Section~\ref{sec:form-secur-with}, we will get back to how a program is instrumented to create an executable configuration.
-% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
@@
 
-\todo[inline]{If target cap. machine is introduced before this section, then revisit/delete this section.}
 % Informally explain how we prevent this using linear capabilities

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Grammar, wording and reference issue.

@@
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, the standard memory and object-like capabilities (sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
+As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
@@
 
-Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
+Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team as well as our own thoughts on the matter.
 As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
@@
 memory location if the value being written is linear.
-A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
+A solution for this problem could be to add separate instructions for moving, storing and loading linear registers at the cost of additional opcode space.
 Adding splice and split will also consume some opcode space.
@@
 In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
+This means that a \texttt{load} of a linear capability will need to behave like an atomic operation similar to a compare-and-swap instruction.
 This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
 When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
+Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle) and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
 If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
@@
 
-Plenty of other papers enforce WBCF at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
 This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
@@
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, \citet{DBLP:journals/corr/abs-1802-00588} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps.
+In follow-up work, \citet{abate_good_components_2018} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Small change to contributions.

@@
 \item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present and use a novel proof technique called \textit{fully-abstract overlay semantics} which we use to prove the properties of \stktokens{}. This includes:
+\item We present a new way to formalise these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
   \begin{itemize}
-  \item defining \srccm{} an overlay semantics for \trgcm{} (Section~\ref{sec:form-secur-with}) and
+  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF guarantees (Section~\ref{sec:form-secur-with})
   \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Shorten.

@@
   \begin{itemize}
-  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF guarantees (Section~\ref{sec:form-secur-with})
+  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
   \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
-  \item using and defining a cross language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
+  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
   \end{itemize}

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Minor edits

--- 2018-07-11 fully-abstract-comp/popl2019/paper.tex
@@
   We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
-  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
-  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
+  The calling convention is based on linear capabilities, a type of capabilities that are prevented from being duplicated by the hardware.
+  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, using what we call a fully abstract overlay semantics.
 \end{abstract}
@@
 \label{sec:introduction}
-\todo[inline]{TODOS ENABLED!!!}
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
@@
 However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
-To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
+To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving, without imposing too much overhead on well-behaved code.
 The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
@@
 % KJAA: This 'sealed' and 'common seal' business is unclear to me.
-For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
+On a recent capability machine called CHERI, for example, they take the form of pairs of capabilities that represent the code and data parts of the closure.
 Each of the two capabilities are sealed with a common seal which make them opaque.
@@
 This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it scales poorly to large amounts of distrusting components because of the need to reserve separate stack space for all components.
-Also in the presence of higher-order values (e.g.
+First, it scales poorly to large amounts of distrusting components, because of the need to reserve separate stack space for all components.
+Also, in the presence of higher-order values (e.g.
 function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
@@
 
-A more scalable approach retains a single stack shared between components.
-Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component and to revoke them after it returns.
+A more scalable approach retains a single stack that is shared between components.
+Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component, and to revoke them after it returns.
 While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
-For example, CHERI offers a form of {\itshape local} capabilities that can only be stored in registers or on the stack but not in other parts of memory.
+For example, CHERI offers a form of {\itshape local} capabilities that, briefly, can only be stored in registers or on the stack, but not in other parts of memory.
 \citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
@@
 % KJAA: Perhaps make it clear by who and when/where?
-These capabilities are non-duplicable enforced by the hardware.
+These capabilities are non-duplicable, enforced by the hardware.
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings particularly without the need for clearing large blocks of memory.
-
-A second contribution of this work is the way in which we formulate these two properties using a technique we call {\itshape fully abstract overlay semantics}.
+The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings, particularly without the need for clearing large blocks of memory.
+
+A second contribution of this work is the way in which we formulate these two properties, using a technique we call {\itshape fully abstract overlay semantics}.
 Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
 Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To accommodate this, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
+To accommodate this, we define a second operational semantics for our target language, with a native well-bracketed stack and primitive ways to do calls and returns.
 This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
@@
 
-Compared to~\citet{skorstengaard_reasoning_2017} that prove LSE and WBCF for a handful of examples,
-this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
+\todo[inline]{In the below, Lars would like to be more precise about the ``in a general way'' below.
+  e.g., by contrasting it to our earlier work where the WBCF and LSE properties were only proved via specific examples.}
+This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
 Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
@@
 
+%\todo[inline]{Add itemized contributions (see comments). Outline?}
 %The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
@@
 \item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present and use a novel proof technique called \textit{fully-abstract overlay semantics} which we use to prove the properties of \stktokens{}. This includes:
+\item We present and use a novel proof technique called fully-abstract overlay semantics which we use to prove the properties of \stktokens{}. This includes:
   \begin{itemize}
   \item defining \srccm{} an overlay semantics for \trgcm{} (Section~\ref{sec:form-secur-with}) and
-  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
-  \item using and defining a cross language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
+  \item proving a full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
+  \item using and defining cross language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
 \end{itemize}
 \end{itemize}
-This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.
+The technical report~\citep{technical_report} contains additional details and proofs that we left out.
 
@@
 \label{sec:cap-mach-w-seal-and-lin}
-% Brief intro to section and this cap machine:
-% Cap machine inspired by CHERI, but with linear capabilities
-\begin{wrapfigure}{r}{0.26\linewidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (rwx) {$\rwx$};
-    \node[main node] (rx) [below right of=rwx] {$\rx$};
-
-    \node[main node] (rw) [below left of=rwx,] {$\rw$};
-    \node[main node] (r) [below right of=rw] {$\readonly$};
-    \node[main node] (0) [below of=r] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-
-    (rw) edge (r)
-    (r) edge (0)
-
-    (rwx) edge (rx)
-
-    (rw) edge (rwx)
-    (r) edge (rx);
-  \end{tikzpicture}
-
-  \caption{Permission hierarchy}
-  \label{fig:perm-hier}
-\end{wrapfigure}
-In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
-For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
-% \lau{Do we have a good reason to assume an infinite address space for this work?
-%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
-
-%%% Capabilities
-The concept of a capability is the cornerstone of any capability machine.
-In its simplest form, a capability is a permission and a range of authority.
-The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
-% Permissions and Range of authority
-The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
-Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
-The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
-% Pointer and linearity
-In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
-The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
-% What is linear capabilities, split/splice
-A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
-The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
-
-%%% Seals
-% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
-
-Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
-%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
-Syntactically, sealed capabilities add two types of capabilities.
-% seals
-A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
-% sealed caps
-A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
-Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
 \begin{figure}[tb]
@@
 \end{figure}
+% Brief intro to section and this cap machine:
+% Cap machine inspired by CHERI, but with linear capabilities
+In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
+For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
+% \lau{Do we have a good reason to assume an infinite address space for this work?
+%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
+
+%%% Capabilities
+The concept of a capability is the cornerstone of any capability machine.
+In its simplest form, a capability is a permission and a range of authority.
+The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
+% Permissions and Range of authority
+The capabilities of \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (as defined in Figure~\ref{fig:target-syntax} along with the rest of the syntax of \trgcm{}).
+Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
+The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
+% Pointer and linearity
+In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
+The linearity $\lin$ is either $\normal$ for traditional capabilities $\linear$ for linear ones.
+% What is linear capabilities, split/splice
+A linear capability is a capability that cannot be duplicated.
+This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source register or memory location is cleared.
+The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
+
+\begin{wrapfigure}{r}{0.28\linewidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (rwx) {$\rwx$};
+    \node[main node] (rx) [below right of=rwx] {$\rx$};
+
+    \node[main node] (rw) [below left of=rwx,] {$\rw$};
+    \node[main node] (r) [below right of=rw] {$\readonly$};
+    \node[main node] (0) [below of=r] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+
+    (rw) edge (r)
+    (r) edge (0)
+
+    (rwx) edge (rx)
+
+    (rw) edge (rwx)
+    (r) edge (rx);
+  \end{tikzpicture}
+
+  \caption{Permission hierarchy}
+  \label{fig:perm-hier}
+\end{wrapfigure}
+
+%%% Seals
+% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
+
+Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
+%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
+Syntactically, sealed capabilities add two types of capabilities.
+% seals
+A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
+% sealed caps
+A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
+Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
 
@@
 Components and linking are defined in Figure~\ref{fig:target-component-and-linking}.
+\todo[inline]{Add better separation between component and linking in figure.}
 
@@
 In Section~\ref{sec:form-secur-with}, we will get back to how a program is instrumented to create an executable configuration.
-% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,\todo[inline]{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
@@
 Some simplifications has been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
-\begin{figure}[tb]
+\begin{figure}[htb]
   \centering
@@
 
+\todo[inline]{If target cap. machine is introduced before this section, then revisit/delete this section.}
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
+As the examples illustrate, the standard memory and object-like capabilities (sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
@@
 Generally speaking, a world is used to model the memory of a language in a logical relation.
-\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
-All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
+\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
+All of the sub-worlds are partial maps from region names (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
 Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
-The region types models linearity as they indicate how the memory can be addressed.
+The region types model linearity as they indicate how the memory can be addressed.
 We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
@@
 The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
-The call stack and free stack is only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
+The call stack and free stack are only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
 Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.
@@
 The future world relation $\future$ is fairly standard.
-It is extensional and requires regions to stay the same type in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.
+It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
+$\spatial$ regions can become $\spatialo$ or $\revoked$.
 
@@
 The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
-The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself which corresponds with our component setup where code memory is read-only.
+The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself which agrees with our component setup where code memory is read-only.
 $H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
-As we will see later, we have two value relations, a trusted and an untrusted, and the value relation used depends on whether the code region governs trusted or untrusted code.
-For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.
+As we will see later, we have two value relations: a trusted and an untrusted.
+The value relation used depends on whether the code region governs trusted or untrusted code.
+For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation includes everything that is in the untrusted value relation in addition to the things that are not safe to pass away in a call, e.g.\ return seals.
 
@@
 It also requires the data return pointer to correspond to one of the stack frames on the call stack.
-Which is expressed by requiring the existence of a spatially owned region in $\Worlds$ that governs memory corresponding to the return pointers range of authority.
+This is expressed by requiring the existence of a spatially owned region in $\Worlds$ that governs memory corresponding to the return pointer's range of authority.
 It must be spatially owned, so it is permitted to address it.
@@
 They both relate capabilities that can safely be used for execution in configurations that satisfy a given world.
-The first one $\mathcal{E}$ is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
+The first one, $\mathcal{E}$, is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
 In our case, the only observable thing on the capability machines is termination.
-The observation relation $\mathcal{O}$ (Figure~\ref{fig:obs-rel}) defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations.
+The observation relation $\mathcal{O}$ defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations.
 However, the only difference between the two approximations is in the observation approximation.}.
-\begin{figure}[htb]
-  \centering
 \begin{align*}
@@
 \end{align*}
-\caption{The observation approximations.}
-\label{fig:obs-rel}
-\end{figure}
 With sealed capabilities in the mix, we also need to be able to say that \texttt{xjmp}s are safe, but a sealed code capability only needs to be safe to execute when it has access to the related data capability.
@@
 \subsubsection{Value relations}
-The purpose of the value relations is two fold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
-As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonable, e.g.\ by treating return seals in a reasonable way.
+The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
+As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave well, e.g.\ by treating return seals in a reasonable way.
 This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.
@@
 Data is always safe, and a piece of data relates to itself.
-A stack pointers is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
+A stack pointer is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
 The stack pointer should only be able to produce safe words, so a stack pointer with read permission can only produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a stack pointer with write permission can at least write safe words.
@@
 An executable memory capability should always point to code which means that there should be a code region governing the memory it points to.
-The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seal is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
+The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seals is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
 % The write condition is called $\stackWriteCond{}$
@@
   For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
-  If $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
+  if\ $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
   and either $[\baddr,\eaddr] \subseteq \ta$ and $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
@@
 \end{lemma}
-The next is a compatibility lemma for plugging that says that given a context and component each related to itself in the component relation, then plugging the context with the component will give an executable configuration in the executable configuration relation.
+The next is a compatibility lemma for plugging that says: given a context and component each related to itself in the component relation, plugging the context with the component will give an executable configuration in the executable configuration relation.
 \begin{lemma}[Compatibility lemma for context plugging]
@@
 
-Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team as well as our own thoughts on the matter.
+Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
 As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
@@
 memory location if the value being written is linear.
-A solution for this problem could be to add separate instructions for moving, storing and loading linear registers at the cost of additional opcode space.
+A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
 Adding splice and split will also consume some opcode space.
@@
 In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability will need to behave like an atomic operation similar to a compare-and-swap instruction.
+This means that a \texttt{load} of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
 This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
 When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle) and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
-If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
+Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
+If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
@@
 The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
-This is a recent, relatively mature capability machine which combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
-For simplicity, we have omitted features of CHERI that were not needed for \stktokens{} (e.g.\ local capabilities, virtual memory).
-
-Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+This is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
+For simplicity, we have omitted features of CHERI that were not needed for modelling \stktokens{} (e.g.\ local capabilities, virtual memory).
+
+Plenty of other papers enforce well-bracketed control flow at a low level, but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
 This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
@@
 Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
-However, its attacker model is different, and its security goals are weaker.
-They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
+However, its attacker model is different and its security goals are weaker.
+They assume an attacker that is not able to execute code, but can overwrite arbitrary data at any time during execution (to model buffer overflows).
 In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
+Also, it only enforces a weak form of control flow correctness, saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
 Such a property ignores temporal properties and seems hard to use for reasoning.
@@
 More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016,juglaret_beyond_2016}.
-Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments), and we do not use per-component stacks or a trusted stack manager but a single shared stack and a decentralized calling convention based on linear capabilities.
-Both prove a secure compilation result from a high-level language which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
+Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
+Both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
 
 CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
-The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation at play.
+The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.
 They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared across component boundaries.
@@
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, \citet{abate_good_components_2018} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
+In follow-up work, \citet{DBLP:journals/corr/abs-1802-00588} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps.
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
+Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler, where these scenarios are not relevant.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex

--- 2018-07-11 fully-abstract-comp/popl2019/paper.tex
@@
 They both relate capabilities that can safely be used for execution in configurations that satisfy a given world.
-The first one $\mathcal{E}$ is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
+The first one, $\mathcal{E}$, is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
 In our case, the only observable thing on the capability machines is termination.
@@
 When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle) and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
-If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
+Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
+If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
@@
 CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
-The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation at play.
+The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.
 They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared across component boundaries.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Shorten.

--- 2018-07-11 fully-abstract-comp/popl2019/paper.tex
@@
 In this section, we present the logical relation.
-Unfortunately due to space considerations, we do not give a comprehensive description of it.
-Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
-Finally for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.
+Unfortunately, for space reasons, we do not give a comprehensive description of it.
+Instead we highlight the parts specific to the logical relation and refer to the \citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
+Finally, for presentation, we will sometimes omit details, like step indexing, that are important for correctness but otherwise not interesting.
 

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Reformulation

--- 2018-07-11 fully-abstract-comp/popl2019/paper.tex
@@
 % \end{itemize}
-The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
+The backbone of the full-abstraction proof is a cross-language, step-indexed, Kripke logical relation.
 It relates the observable behavior of \srccm{} to the observable behavior of \trgcm{}.
@@
 The value relation used depends on whether the code region governs trusted or untrusted code.
-For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation includes everything that is in the untrusted value relation in addition to the things that are not safe to pass away in a call, e.g.\ return seals.
+For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and %the trusted value relation contains the untrusted value relation and the things that are not safe to pass away in a call, e.g.\ return seals.
+the untrusted value relation is subsumed in the trusted value relation that also includes the things that are not safe to pass away in a call, e.g.\ return seals.
 
@@
 The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
-As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave well, e.g.\ by treating return seals in a reasonable way.
+As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonable, e.g.\ by treating return seals in a reasonable way.
 This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Shorten, rephrase.

@@
 \label{subsec:worlds}
-Generally speaking, a world is used to model the memory of a language in a logical relation.
-\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
-All of the sub-worlds are partial maps from region names (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
+Generally speaking, a world is used in a logical relation to model assumptions about memory.
+\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
+All of the sub-worlds are partial maps from $\RegionName$ (modelled as natural numbers) to regions.
+The regions differ for each of the sub-worlds based on what they model.
 Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
-The region types model linearity as they indicate how the memory can be addressed.
+The region types models linearity as they indicate how the memory can be referenced.
 We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
-The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
+The $\pure$ regions can be referenced by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
 To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
-To this end, we have the relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
+To this end, we have the partial operation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except that they have disjoint ownership: $\pure$ regions can be present in both but a $\spatialo$ region can only be present in one of both worlds.
+Such a region must then be $\spatial$ in the other world.
 %We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
-The unaddressable $\spatial$ region makes sure that the memory it represents is modelled.
-Linear capabilities can be seen as revocable which is modelled by the $\revoked$ region.
-Finally, the $\pure$ region also contains a seal interpretation function.
-The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
+This is important because the logical relation will require that valid memories are backed by a suitable memory segment, even when the memory does not own them (for example, when a linear capability for it is in one of the registers).
+Revocation of linear capabilities is supported in a certain way using the $\revoked$ region.
+Finally, $\pure$ regions can also claim ownership of a seal by defining a seal interpretation function at this seal.
+The seal interpretation is a world-indexed relation that must hold for all sealables signed by this seal.
 We split the regions in two categories:
@@
 The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
-The call stack and free stack are only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
-Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.
+The call stack and free stack are only ever addressed by linear capabilities, so their sub-worlds only allow $\Regions$.
+Regions in $\Worlds$ correspond to a stack frame, so they contain the return address for this frame.
 
@@
 % \end{lemma}
-The future world relation $\future$ is fairly standard.
-It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
-$\spatial$ regions can become $\spatialo$ or $\revoked$.
+The future world relation $\future$ is fairly simple.
+It is extensional and requires regions to stay the same type in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.
 
@@
 The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
-The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself which agrees with our component setup where code memory is read-only.
+The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself, i.e.\ our component's code memory must not be modified.
 $H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
-As we will see later, we have two value relations: a trusted and an untrusted.
-The value relation used depends on whether the code region governs trusted or untrusted code.
-For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and %the trusted value relation contains the untrusted value relation and the things that are not safe to pass away in a call, e.g.\ return seals.
-the untrusted value relation is subsumed in the trusted value relation that also includes the things that are not safe to pass away in a call, e.g.\ return seals.
+As we will see later, we have two value relations, a trusted and an untrusted, and the value relation used depends on whether the code region governs trusted or untrusted code.
+For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation additionally includes things that trusted components are allowed to have but not pass away, e.g.\ return seals.
 
@@
 The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigclos$ can seal.
-The return seals can only be used to seal return pointers that corresponds to the return from an call in the code of the region.
+The return seals can only be used to seal return pointers that corresponds to the return from a call in the code of the region.
 Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
@@
 Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
-Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function related data return pointers to linear capabilities.
-It also requires the data return pointer to correspond to one of the stack frames on the call stack.
-This is expressed by requiring the existence of a spatially owned region in $\Worlds$ that governs memory corresponding to the return pointer's range of authority.
-It must be spatially owned, so it is permitted to address it.
-A region in $\Worlds$ is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
+Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function relates data return pointers to linear capabilities.
+It also requires the data return pointer to correspond to a $\spatialo$ region for one of the stack frames on the call stack.
+The return address that is part of this region in $\Worlds$, must be the return address of the call that the return seal corresponds to.
 \begin{multline*}

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Making everything stick withing margins.
Disable todo's
preamble updated with documentclass that shows header (not active)
typo

@@
 \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
+%\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 
-\usepackage{todonotes}
+\usepackage[disable]{todonotes}
 
@@
    \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
-    \aendbnf \in & \Addr \cup \{\infty \} & & & & \sealendbnf \in & \Seal \cup \{\infty \}& &\\
+    \aendbnf \in & & & \Addr \cup \{\infty \} & & \sealendbnf \in & & & \Seal \cup \{\infty \}\\
     \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
@@
 \]
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.8cm}|}
+  \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.7cm}|}
     \hline
@@
   \end{mathpar}
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.8cm}<{$}|>{\raggedright\arraybackslash}p{6.5cm}|}
+  \begin{tabular}{|>{$}c<{$}|>{$}p{3.8cm}<{$}|>{\raggedright\arraybackslash}p{6.4cm}|}
     \hline
@@
 Unfortunately, for space reasons, we do not give a comprehensive description of it.
-Instead we highlight the parts specific to the logical relation and refer to the \citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
+Instead we highlight the parts specific to the logical relation and refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
 Finally, for presentation, we will sometimes omit details, like step indexing, that are important for correctness but otherwise not interesting.

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex
Changes.

@@
 Generally speaking, a world is used in a logical relation to model assumptions about memory.
-\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
+\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
 All of the sub-worlds are partial maps from $\RegionName$ (modelled as natural numbers) to regions.
@@
 Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
-The region types models linearity as they indicate how the memory can be referenced.
+The region types model linearity as they indicate how the memory can be referenced.
 We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
@@
 To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
-To this end, we have the partial operation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except that they have disjoint ownership: $\pure$ regions can be present in both but a $\spatialo$ region can only be present in one of both worlds.
+To this end, we have the partial operation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except that they have disjoint ownership: $\pure$ regions can be present in both worlds, but a $\spatialo$ region can only be present in one.
 Such a region must then be $\spatial$ in the other world.
@@
 Revocation of linear capabilities is supported in a certain way using the $\revoked$ region.
-Finally, $\pure$ regions can also claim ownership of a seal by defining a seal interpretation function at this seal.
+Finally, $\pure$ regions can also claim ownership of a seal by defining a seal interpretation for it.
 The seal interpretation is a world-indexed relation that must hold for all sealables signed by this seal.
@@
 % \end{lemma}
-The future world relation $\future$ is fairly simple.
-It is extensional and requires regions to stay the same type in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.
+The future world relation $\future$ is fairly standard.
+It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
+$\spatial$ regions can become $\spatialo$ or $\revoked$.
 
@@
 $H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
-As we will see later, we have two value relations, a trusted and an untrusted, and the value relation used depends on whether the code region governs trusted or untrusted code.
+As we will see later, we have two value relations: a trusted and an untrusted.
+The value relation used depends on whether the code region governs trusted or untrusted code.
 For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation additionally includes things that trusted components are allowed to have but not pass away, e.g.\ return seals.
@@
 It also requires the data return pointer to correspond to a $\spatialo$ region for one of the stack frames on the call stack.
-The return address that is part of this region in $\Worlds$, must be the return address of the call that the return seal corresponds to.
+The return address that is part of this region in $\Worlds$, must be the return address of the call that the return seal is affiliated with.
 \begin{multline*}

### 2018-07-12 fully-abstract-comp/popl2019/paper.tex
Small changes to section 5.

@@
 \renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
-\renewcommand{\Worlds}{\World_{\mathrm{call\ stack}}}
+\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
+\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
 
@@
   The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
-  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
+  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalize and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
 \end{abstract}
@@
 \item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present a new way to formalise these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
+\item We present a new way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
   \begin{itemize}
@@
 Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
-The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
+The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$) ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
 % Pointer and linearity
@@
 Unfortunately, for space reasons, we do not give a comprehensive description of it.
-Instead we highlight the parts specific to the logical relation and refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
+Instead we highlight specific parts to the logical relation and refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
 Finally, for presentation, we will sometimes omit details, like step indexing, that are important for correctness but otherwise not interesting.
@@
 The region types model linearity as they indicate how the memory can be referenced.
-We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
-The $\pure$ regions can be referenced by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
+We have four types of regions: $\pure$, $\spatial$, $\spatialo$, and $\revoked$.
+The $\pure$ regions can be referenced by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ and $\revoked$ regions by nothing.
 To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
@@
 %We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
-This is important because the logical relation will require that valid memories are backed by a suitable memory segment, even when the memory does not own them (for example, when a linear capability for it is in one of the registers).
+This is important because the logical relation will require that valid memories are backed by a suitable memory segment even when the memory does not own them (for example, when a linear capability for it is in one of the registers).
 Revocation of linear capabilities is supported in a certain way using the $\revoked$ region.
@@
 
-Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$.
+Define worlds as $\World = \Worldh \times \Worlds \times \Worldfs$.
 They are recursive as they contain world indexed relations.
@@
 It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
-$\spatial$ regions can become $\spatialo$ or $\revoked$.
+The $\spatial$ regions can become $\spatialo$ or $\revoked$.
 
@@
 
-The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigclos$ can seal.
+The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigcloss$ can seal.
 The return seals can only be used to seal return pointers that corresponds to the return from a call in the code of the region.
 Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
-The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
-The code return pointer corresponds to a normal capability which means that it can be reused but always to return from the same call.
+The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (Figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
+The code return pointer corresponds to a normal capability which means that it can be reused but only to return from the same call.
 Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
 Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function relates data return pointers to linear capabilities.
-It also requires the data return pointer to correspond to a $\spatialo$ region for one of the stack frames on the call stack.
-The return address that is part of this region in $\Worlds$, must be the return address of the call that the return seal is affiliated with.
+It also requires the data return pointer to refer to a $\spatialo$ region for one of the stack frames on the call stack.
+The return address of this region in $\Worlds$ must be the return address of the call that the return seal is affiliated with.
 \begin{multline*}
@@
 The closure seals in untrusted code should therefore only be used to seal things from the untrusted value relation, i.e.\ things that are safe to transfer to an untrusted component.
-The trusted closure seals can by used to seal things from the untrusted value relation.
+The trusted closure seals can be used to seal things from the untrusted value relation.
 This allows trusted seals to seal capabilities for the trusted code which is necessary in order to create a new closure.
@@
 The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
-This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that corresponds with the view \srccm{} provides on the memory.
-Each of the three partitions of $\ms_T$ must be related to the three parts of \srccm{} memory under the relevant sub-worlds of $W$.
+This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that correspond with the view \srccm{} provides on the memory.
+Each of the three partitions of $\ms_T$ must be related to one of the three parts of \srccm{} memory under the relevant sub-world of $W$.
 The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
-The free stack $\ms_\stk$ relates to the free stack partition of $\ms_T$ in the standard way, but with respect to the free stack world.
+The free stack $\ms_\stk$ also relates to the free stack partition of $\ms_T$ in the standard way but with respect to the free stack world.
 %Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
-The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts relating each of the stack frames.
-Further, as each of the stack frames correspond to a slot in the return order, none of the stack frames should be empty.
-It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved in the actual memory.
+The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts related to each of the stack frames.
+Further, as each of the stack frames corresponds to a slot in the return order none of the stack frames should be empty.
+It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved by the partitions of $\ms_T$.
 
@@
 The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
-As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonable, e.g.\ by treating return seals in a reasonable way.
-This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.
-This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and untrusted value relation.
+As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonably, e.g.\ by handling return seals in a safe way.
+This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and want to guarantee LSE and WBCF against.
+This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and an untrusted value relation.
 The two value relations are sketched in Figure~\ref{fig:value-relation}.
 
-The untrusted value relation relates words that can safely be given to untrusted code.
+The untrusted value relation relates words from the two machines that can safely be given to untrusted code.
 Data is always safe, and a piece of data relates to itself.
 A stack pointer is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
-The stack pointer should only be able to produce safe words, so a stack pointer with read permission can only produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a stack pointer with write permission can at least write safe words.
-Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack part of the world.
-Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read-write permission.
+A safe stack pointer with read permission should only be able to produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a safe stack pointer with write permission can at least write safe words.
+Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack sub-world.
+Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read (and write) permission.
 The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
-An executable memory capability should always point to code which means that there should be a code region governing the memory it points to.
+An executable memory capability should always point to code which means that there should be a code region governing the memory it references.
 The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seals is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
@@
 Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
-The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\vsc_S,\vsc_T)$ must be in the seal interpretation function for $\sigma$.
+The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\src{\vsc_S},\vsc_T)$ must be in the seal interpretation function for $\sigma$.
 Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.
@@
 Trusted code uses the return seals and the closure seals (in a responsible way), and therefore sets of seals with trusted return seals and closure seals are in the trusted value relation.
-In order for trusted code to even execute, it needs to use an executable memory capability for the trusted code, so such capabilities are in the untrusted value relation.\footnote{Note that code capabilities for trusted code is not allowed in the untrusted value relation as it could be used to read return seals that are not in the untrusted value relation.}
+In order for trusted code to even execute, it needs to use an executable memory capability for the trusted code, so such capabilities are in the trusted value relation.\footnote{Note that code capabilities for trusted code is not allowed in the untrusted value relation as it could be used to read return seals that are not in the untrusted value relation.}
 
@@
   \label{thm:ftlr}
-  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
-  if\ $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
-  and either $[\baddr,\eaddr] \subseteq \ta$ and $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
+  If\ $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
+  and either $[\baddr,\eaddr] \subseteq \ta$ and\\ $\src{((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
 $[\baddr,\eaddr] \mathrel{\#} \ta$,
@@
   \[
-    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr), ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
+    \npair{\left(\src{((\rx,\normal),\baddr,\eaddr,\aaddr)}, ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
   \]
 \end{lemma}
-Roughly speaking, the FTLR says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
+Roughly speaking, the FTLR says: any executable capability used for execution in a safe environment won't break the capability machine guarantees.
 This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ entails, and for trusted code the semantic reasonability condition must also be satisfied.
-The proof of the FTLR is done by a nested induction where each instruction that could be executed is considered.
+The proof of the FTLR is done by a nested induction where each instruction that could be executed is proven safe.
 The lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, but it is crucial to the proofs of the lemmas in that section.
@@
 \label{subsec:proof-sketch}
-The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward.
+The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward manner.
 The component relation relates components to themselves which makes sense as well-formed components should contain no \srccm{} specific constructs.
@@
   If $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$ and either
-    $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component; or
+    $\dom(\comp.\mscode) \subseteq \ta$ and $\src{\comp}$ is a reasonable component; or
     $\dom(\comp.\mscode) \mathrel{\#} \ta$,
   then there exists a $W$ such that
-  $\npair{(\comp,\comp)} \in \lrcomp(W)$
+  $\npair{(\src{\comp},\comp)} \in \lrcomp(W)$
 \end{lemma}
@@
   If $\npair{\stpair{\context}{\context}}\in\lrcomp(W_1)$ and $\npair{\stpair{\comp}{\comp}} \in \lrcomp(W_2)$, then
-  $\npair[n']{(\plug{\context_S}{\comp_S},\plug{\context_T}{\comp_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
+  $\npair[n']{(\plug{\src{\context_S}}{\src{\comp_S}},\plug{\context_T}{\comp_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
 \end{lemma}
@@
  $i \leq n$ and
-and $\Phi_S \sterm[i]{\gc}$,
+and $\src{\Phi_S} \sterm[i]{\gc}$,
   then $\Phi_T\term$.
@@
 and $\Phi_T \term[i]$,
-  then $\Phi_S\sterm{\gc}$.
+  then $\src{\Phi_S}\sterm{\gc}$.
 \end{lemma}
@@
   The proof of both directions are very similar, so we only show the right direction.
-  W.l.o.g.\ it suffices to show one direction of the contextual equivalence, so we assume ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
+  We need to show the \trgcm{} contextual equivalence, so w.l.o.g\ assume ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
   The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.
-  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that the trusted components $\comp_1$ and $\comp_2$ are well-formed and reasonable.
-  Further, by $\tconeq$ we may assume the context $\context$ is well-formed.
-  We prove arrow (1) in the figure by using the mentioned assumptions about $\comp_1$ and $\context$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
+  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that the trusted components $\src{\comp_1}$ and $\src{\comp_2}$ are well-formed and reasonable.
+  Further, by $\tconeq$ we may assume the context $\trg{\context}$ is well-formed.
+  We prove arrow (1) in the figure by using the mentioned assumptions about $\src{\comp_1}$ and $\trg{\context}$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
   Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$, so by the assumption that $\comp_1$ and $\comp_2$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$, i.e.\ arrow (2) in the figure.
-  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use the assumption that $\comp_2$ is well-formed and reasonable and that $\context$ is well-formed.
+  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use the assumption that $\src{\comp_2}$ is well-formed and reasonable and that $\trg{\context}$ is well-formed.
 \end{proof}

### 2018-07-17 fully-abstract-comp/popl2019/paper.tex
todos for camera ready and added missing captions.

@@
 \todo[inline]{TODOS ENABLED!!!}
+\todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
+\todo[inline]{Lau: for camera ready version take a look at the coloring in the paper}
+\todo[inlint]{Lau: for camera ready version if TOPLAS paper accepted, update the reference in LR section. }
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
@@
 % \end{definition}
-  \caption{Components and linking of components}
+  \caption{Components and linking of components.}
   \label{fig:target-component-and-linking}
@@
 
-  \caption{Possible ways to abuse stack and return capabilities}
+  \caption{Possible ways to abuse stack and return capabilities.}
   \label{fig:stack-ret-ptr-abuse}
@@
   \end{subfigure}
-  \caption{}
+  \caption{Abuse of stack and return capabilities prevention.}
 \end{figure}
@@
   \end{subfigure}
-  \caption{}
+  \caption{Partial return token used to return out of order.}
   \label{fig:stk-base-abuse}
@@
 \end{figure}
+
+\todo[inline]{Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.}
 

### 2018-07-17 fully-abstract-comp/technical-report.tex
Proposal for padding of stack (yet to be checked).
Small update to previous fix and one proof.
Removed TODO
Compilation issue fixed.
Removed the rest of a todo that was removed previously.
Added T_A to well-formedness judg., fixed comp jdg., and fixed full-abstraction pf.
T_A was added to main judgement and comp judgement.
Now the call case of the comp judgement puts no restrictions on calls outside of
ta.
The full-abstraction proof needed to be fixed with a lemma that allows the
trusted addresses in the well-formedness judgement to be increased to the ones
of the program that is plugged into

@@
 \end{align*}
-\lau{(important) TODO 05-07-2018: It seems that there may be an issue related to splicing stk-ptr with a linear capability.
-  This would be okay on trg machine, but fail on src machine.
-  It may possibly be fixed simply by padding the stack with un-addressable memory (like we did with code memory).}
 Two important points about $\tsplice{}{}{}$ related to the calling convention: (1) Splice fails if two capabilities are not adjacent.
@@
 \clearpage
-We define inductively when a component is valid ($\vdash \var{comp}$) by the below inference rules:
+We define inductively when a component is valid ($\ta \vdash \var{comp}$) by the below inference rules:
 \begin{mathpar}
@@
   }{
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
   }
@@
     \mscode(a) \in \ints\\
-    \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
-  }{
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a
+{    \begin{multlined}
+      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
+      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
+    \end{multlined}
+}  }{
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
   }
@@
     \forall a \in \dom(\mscode) \ldotp 
-    \sigrets,d_\sigma(a),\sigcloss \vdash_{\mathrm{comp-code}} \mscode,a\\
+    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
     \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
   }{
-    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode
+    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
   }
@@
     \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
-    \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
+    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
     \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
@@
     \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}\\
-    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
+    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} & (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
     \dom(\msdata) \mathrel{\#} \ta
   }{
-    \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
   }
@@
     \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
-    \vdash \var{comp}_0 & (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
+    \ta \vdash \var{comp}_0 & (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
   }{
-    \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
+    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
   }
@@
       \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \uplus \ms_\var{frame} \\
-      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
+      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) &
+      [\baddr_\stk-1,\eaddr_\stk+1]\mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
       \overline{\var{import}} = \emptyset
@@
 \end{definition}
+
+\begin{lemma}
+  \label{lem:context-trusted-addr-ampl}
+  For components $\context$ and $\comp$,
+  if
+  \begin{itemize}
+  \item $\emptyset \vdash \context$
+  \item $\dom(\comp.\mscode) \vdash \comp$
+  \item $\plug{\context}{\comp}$ is defined
+  \end{itemize}
+  Then
+  \[
+    \dom(\comp.\mscode) \vdash \context
+  \]
+\end{lemma}
+\begin{proof}
+  Follows by definition.
+\end{proof}
 
@@
       \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
-
+      \quad \dom(\ms_{\var{T,stack}} \uplus \var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
+      \quad \{\baddr_\stk -1,\eaddr_\stk + 1\} \in \dom(\ms_{T,f}) \wedge \\
       \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
@@
       \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
-      \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
+      \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \code \wedge\\
       \forall a \in \dom(\code)\ldotp\\
@@
   \begin{equation*}
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss \vdash_{\mathrm{comp-code}} w
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta\vdash_{\mathrm{comp-code}} w
   \end{equation*}
@@
   \item $\mscode(a) \in \ints$ and
-    \begin{equation*}
-      \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
-    \end{equation*}:
+    \begin{multline*}
+      ([a \cdots a + \calllen-1] \subseteq \ta \wedge\mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
+    \end{multline*}:
 
@@
   \item $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$
-  \item $\var{comp}$ is a well-formed component, i.e. $\vdash \var{comp}$
+  \item $\var{comp}$ is a well-formed component, i.e. $\ta \vdash \var{comp}$
   \item One of the following holds:
@@
   Note that confusingly, $\var{comp}.\mscode = \mscode\uplus\mspad$, so the assumption about $\dom(\var{comp}.\mscode)$ should be interpreted properly.
-  We then know from $\vdash \var{comp}$ that 
+  We then know from $\ta \vdash \var{comp}$ that 
   \begin{itemize}
@@
   \item $\mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]$
-  \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode$
+  \item $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$
   \item $\exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}$
@@
       \item $([\baddr-1,\eaddr+1] \wedge \trust = \trusted) \vee ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)$: by assumption and choice of $\trust$.
-      \item $\sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code$: by assumption.
+      \item $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \code$: by assumption.
       \item $\mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]$:
@@
   \item $\mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk}$
-  \item $[\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata))$
+  \item $[\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk )$
+  \item $[\baddr_\stk-1,\eaddr_\stk+1] \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata))$
   \item $\overline{\var{import}} = \emptyset$
@@
   \begin{equation*}
-    \forall \src{\context} \ldotp \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
+    \forall \src{\context} \ldotp \emptyset \vdash \src{\context} \Rightarrow \src{\plug{\context}{\var{comp}_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\var{comp}_2} \sterm{\ta[,2],\stkb_2}}
   \end{equation*}
@@
   \begin{equation*}
-    \forall \trg{\context} \ldotp \vdash \trg{\context} \Rightarrow \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
+    \forall \trg{\context} \ldotp \emptyset \vdash \trg{\context} \Rightarrow \trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow \trg{\plug{\context}{\var{comp}_2} \term}
   \end{equation*}
@@
   \label{thm:full-abstraction}
-  For reasonable, well-formed components $\var{comp}_1$ and $\var{comp}_2$, we have
+  For reasonable, well-formed components $\var{comp}_1$ and $\var{comp}_2$ (with respect to $\ta[,i] = \dom(\var{comp}_i.\mscode)$, respectively), we have
   \begin{gather*}
@@
 
-    Take a $\src{\context}$ such that $\vdash \src{\context}$, take $\src{\ta[,i]}
+    Take a $\src{\context}$ such that $\emptyset \vdash \src{\context}$, take $\src{\ta[,i]}
     = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and
@@
     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\preceq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} and Lemma~\ref{lem:context-trusted-addr-ampl} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\preceq,\gc_1](W_1\uplus W_1')$
@@
     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\succeq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} and Lemma~\ref{lem:context-trusted-addr-ampl} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\succeq,\gc_2](W_2\uplus W_2')$
@@
 
-    Take a $\trg{\context}$ such that $\vdash \trg{\context}$ and we will prove that
+    Take a $\trg{\context}$ such that $\emptyset \vdash \trg{\context}$ and we will prove that
     $\trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow
@@
     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} and Lemma~\ref{lem:context-trusted-addr-ampl} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
@@
     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-    It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
+    It also follows from the same Lemma~\ref{lem:ftlr-comps} and Lemma~\ref{lem:context-trusted-addr-ampl} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$

### 2018-07-17 fully-abstract-comp/technical-report.tex
Removed todos.

@@
 \end{definition}
-\lau{07-07-2018: TODO, consider if the well-formedness judgement should take the trusted addresses into account, so the trusted component specific things only are required for trusted addresses.
-It already uses $\ta$, so it already ought to be part of the judgement.
-}
-\lau{07-07-2018: TODO: why don't we permit closure seals in data memory? That seems like an oversight.}
 

### 2018-07-18 fully-abstract-comp/technical-report.tex
typo
Fixed todo.
Fixed TODO about missing connection between ret/clos seals and the global ones
The issue was in a lemma that may not have been fixed after introduction of
global seals.

@@
 \end{definition}
-\lau{07-07-2018: TODO why don't the above definition case about the stack base?}
+\lau{07-07-2018: TODO why don't the above definition care about the stack base?}
 
@@
 Memory satisfaction for new worlds:
-\lau{09-07-2018: TODO: add $\gc$ to memsat relations.}
 \[
@@
   \begin{itemize}
+  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
   \item $\pwheap(r_{\mathrm{code}}) = \codereg{\sigrets,\sigcloss,\mscode,\gc}$
@@
 
-\lau{09-07-2018: TODO: It seems very odd to me that the return and closure seals in $\gc$ are independent of the seals in the component.
-I would suspect that in the trusted case, $\gsigrets,\gsigcloss$ is a suberset of the seals in the component (or something like that).
-}
 \begin{lemma}[FTLR for components]
@@
     \end{itemize}
+  \item $\sigrets \subseteq \gsigrets$ and $\sigcloss \subseteq \gsigcloss$
   \end{itemize}
@@
       \item $\dom(\code) = [\baddr,\eaddr]$: by assumption.
-      \item $([\baddr-1,\eaddr+1] \wedge \trust = \trusted) \vee ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)$: by assumption and choice of $\trust$.
+      \item $([\baddr-1,\eaddr+1]\subseteq \ta \wedge \trust = \trusted) \vee ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted)$: by assumption and choice of $\trust$.
       \item $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \code$: by assumption.

### 2018-10-07 fully-abstract-comp/popl2019/paper.tex
Dealt with some of reviewer A's and B's comments

@@
 There is, however, no guarantee that untrusted assembly code respects this intended usage.
-Particularly, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.
+In particular, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.
 
@@
 To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
-A first approach uses separate stacks for distrusting components, and a central, trusted stack manager component that mediates cross-component invocations.
+A first approach uses separate stacks for mutually distrusting components, and a central, trusted stack manager component that mediates cross-component invocations.
 This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it scales poorly to large amounts of distrusting components because of the need to reserve separate stack space for all components.
+First, it scales poorly to large amounts of components because of the need to reserve separate stack space for all components.
 Also in the presence of higher-order values (e.g.
@@
 % KJAA: Perhaps make it clear by who and when/where?
-These capabilities are non-duplicable enforced by the hardware.
+The hardware prevents these capabilities from being duplicated.
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings particularly without the need for clearing large blocks of memory.
+The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings in particular without the need for clearing large blocks of memory.
 
@@
               \and
-              \inference{ \neg (\Phi \step \Phi') \and \Phi' \neq \failed}
+              \inference{ \forall \Phi' \ldotp \neg (\Phi \step \Phi') \wedge \Phi' \neq \failed}
                         {\Phi \step \failed }
@@
     \hline
-    \move{r}{\rn}                                & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\rn}$ \\
+    \move{r}{\rn}                                & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\Phi(\rn)}$ \\
     \hline
@@
     \hline
-    \jmp{r}    &\Phi\updReg{r,\pcreg}{w,\Phi(r)} & $w = \linCons{\Phi(r)}$\\
+    \jmp{r}    &\Phi\arraycolsep=0pt\array[t]{l}\updReg{r}{w}\\\updReg{\pcreg}{\Phi(r)}\endarray & $w = \linCons{\Phi(r)}$\\
     \hline
@@
     \hline
-    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_1)}$\\
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_3)}$\\
     \hline
@@
 The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability instructions simply inspect the properties of capabilities: type (\texttt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
+The simplest of the capability instructions inspect the properties of capabilities: type (\texttt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
 The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
@@
 In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
-If both things are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relatively to $\Phi$.
-The interpretation of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
+If both things are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
+The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
 In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
@@
 This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
-Opaque sealed capabilities encapsulate a security domain's local state and authority, and it only becomes accessible again when control is transferred to the security domain.
+Opaque sealed capabilities encapsulate a security domain's local state and authority, and they only become accessible again when control is transferred to the security domain.
 Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one closure with the environment of another.
@@
 
-At this time, the executable configuration is our only way to specify the state of the capability machine.
-We would, however, much prefer a notion of programs that only describes the essential parts of an entity on the machine and only later instrument the program to make it into an executable configuration.
-In later sections, we describe different entities on the capability machine that interact with each other.
-In order for them to interact, we need to be able to link them together, and to this end we introduce a notion of components.
-A component is basically a program with open ends in the form of imports that need to be satisfied.
-A component also has exports allowing other components to link with it.
-We define the notion of a component in Figure~\ref{fig:target-component-and-linking}.
+The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
+To make this clearer, we introduce notions of components and programs from which we construct executable configurations.
+A component is basically a program with entry points in the form of imports that need to be linked.
+A component has exports that can satisfy the imports of other components.
+The component definition is in Figure~\ref{fig:target-component-and-linking}.
 % Structure
-A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses.
+A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses (addresses governed by a linear capability).
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
@@
 %base = library
-A base component can be seen as library implementations because it does not specify a main entry point of execution.
+A base component has no main entry point, so it needs to be invoked by another component to run. For this reason, it can be seen as a library implementation.
 A component is either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
@@
 Libraries can be linked together and incomplete programs can be linked with libraries, but two incomplete programs cannot be linked with each other.
-Two components can be linked when their memories, seals, and linear addresses disjoint.
+Two components can be linked when their memories, seals, and linear addresses are disjoint.
 They are combined by taking the union of each of their constituents.
@@
 The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
-Components and linking are defined in Figure~\ref{fig:target-component-and-linking}.
 
@@
 
-Some simplifications has been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
+Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
 \begin{figure}[tb]
@@
   \inference{
-    \comp_1 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
-    \comp_2 = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
-    \comp_3 = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
+    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
+    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
+    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
     \mscode[3] = \mscode[1] \uplus \mscode[2] \\
@@
   } {
-    \comp_3 = \comp_1 \bowtie \comp_2
+    \comp_0'' = \comp_0 \bowtie \comp_0'
   }
@@
 %       c_{\mathrm{main},c} = \sealed{\sigma_1, c_{\mathrm{main},c}'} &
-%       c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'} &
+%       c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'} &comp\
 %       \sigma_1 = \sigma_2\\
@@
 At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
-However, this does not make sure that the caller's stack frame is on top of the call stack.
+However, this does not ensure that the caller's stack frame is on top of the call stack.
 The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
@@
 All call points need to be associated with a unique seal (a return seal) that is only used to seal the capabilities in the return pair for that particular call point.
-The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack which would break WBCF.
+The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack, which would break WBCF.
 For similar reasons, we cannot allow return seals to be used to seal closures.
@@
 }
-  \caption{The instructions that constitutes a call.
-The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
+\caption{
+  The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
+  Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
+  Instructions 1-3: ensure that the stack is non-empty.
+  Instructions 4-5: split stack pointer in two (stack pointer for callee done).
+  Instructions 6-9: Load the seal designated for this call.
+  Instruction 10: Seal capability for private stack to create data return pointer.
+  Instructions 11-13: Seal capability for return point to create code return pointer.
+  Instructions 14-15: Clean up temporary register and jump.
+  Instructions 16-26: The return code.
+  Instructions 16-23: check the stack base.
+  Instruction 24: join the stack pointer.
+  Instructions 25-26: pop 42 from the stack.
+%The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
 The variable $\offsigma$ is the offset in the set of seals.
@@
 Line 19: $5$, offset to fail.
-Line 21: offset to address after fail.}
+Line 21: offset to address after fail.
+}
   \label{fig:call-code}
@@
 Our formal results are complicated by the fact that they only hold for well-behaved components that respect the basic rules of the calling convention, and also on a sane initial configuration of the system.
-For example, the system should be setup such that seals that components use for constructing return pointers should not be shared with other components.
+For example, the system should be set up such that seals that components use for constructing return pointers are not shared with other components.
 We envision distributing seals as a job for the linker, so this means our results depend on a linker to do this properly.
-As another example, a seal that is used to construct a return pointer, can be used again to construct other return pointers for the same return point, but a different seal must be used for others.
+As another example, a seal that is used to construct a return pointer can be used again to construct other return pointers for the same return point, but a different seal must be used for others.
 Such seals should also never be passed to other components.
@@
 Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to on the underlying machine.
-Return pointers are generated by calls from the capabilities they correspond to, and they are turned back to the capabilities they correspond to upon return.
-Even though the return pointers both correspond to capabilities with read permission, neither can be used for reading. 
-Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack memory that is encapsulated in a stack frame which would break LSE.
+On \srccm{}, return pointers are generated by calls from the capabilities they correspond to on \trgcm{}, and they are turned back to the capabilities they correspond to upon return.
+% Even though the return pointers both correspond to capabilities with read permission, neither can be used for reading. 
+% Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack memory that is encapsulated in a stack frame which would break LSE.
 
@@
 The stack frame consists of the local stack (addresses $[\baddr_\stk,\aaddr_\stk-1]$) as well as the address where execution is supposed to continue on return.
-The return address in the stack frame is an important piece of meta data as it connects the stack frame to the return point and thus allows us on return to check whether we returned to the right place.
+The return address in the stack frame is an important piece of metadata as it connects the stack frame to the return point and thus allows us on return to check whether we returned to the right place.
 The return pointers are sealed with the seal dictated by $\offpc$ and $\offsigma$ and placed in registers $\rretc$ and $\rretd$, so the callee can use them to return.
@@
 The return pair can only be used to return with.
-They only exists on \srccm{} in sealed capabilities, so it is only natural that \texttt{xjmp} takes care of returning.
+They only exist on \srccm{} in sealed capabilities, so it is only natural that \texttt{xjmp} takes care of returning.
 To this end, we update the semantics of \texttt{xjmp} by redefining $\xjmpres{}$.
@@
 \srccm{} supports tail calls.
-A tail call is a call from a caller that is done executing and thus don't need to be returned to.
+A tail call is a call from a caller that is done executing and thus doesn't need to be returned to.
 This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in call.
@@
 In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
-We more precisely expect four things of a reasonable component: (1) It checks the stack base address before performing a call.
+More precisely, we expect four things of a reasonable component: (1) It checks the stack base address before performing a call.
 As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
@@
 \label{subsec:proof-sketch}
-The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward manner.
+The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straightforward manner.
 The component relation relates components to themselves which makes sense as well-formed components should contain no \srccm{} specific constructs.

### 2018-10-08 fully-abstract-comp/popl2019/paper.tex
Addressed some comments from reviewer C

@@
 A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
-Sealable capabilities $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
+Sealables $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
 \begin{figure}[tb]
@@
 Because the stack capability is linear the caller knows that the capability for their local stack frame cannot have an alias.
-This means that an adversary will need the stack capability the caller produced in order to access their local data.
+This means that an adversary would need the stack capability the caller produced
+in order to access the caller's local data.
 While the caller gives the capability to the adversary, it is only after the caller seals it which makes the capability opaque.
@@
 The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
-At this point (Figure~\ref{fig:stack-base-abuse-b})), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
+At this point (Figure~\ref{fig:stack-base-abuse-b}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
 For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In \stktokens{}, the base address of the stack is fixed as a compile-time constant.
+In \stktokens{}, the base address of the stack capability is fixed as a
+compile-time constant (Note: the stack grows downwards, so the base address of
+the stack capability is the top-most address of the stack). 
 The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
@@
 Only instructions at these addresses are allowed to perform calls that push stack frames to the call stack, i.e.\ calls for which \srccm{} provides native LSE and WBCF guarantees.
-The constant $\ta$ is passed around as a parameter of the \trgcm{} step relation.
+The constant $\ta$ is passed around as a parameter of the \srccm{} step relation.
 Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
@@
 This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
-The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ({\footnotesize $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is $26$, i.e.\ the number of instructions).
+The sequence of instructions have to correspond to a call, i.e.\ the
+instructions in Figure~\ref{fig:call-code} ({\footnotesize  $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is always $26$, i.e.\ the number of instructions).
 Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.
@@
 If a component is not trusted, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
-The code memory may contain sets of seals, but only the seals in $\sigrets$ and $\sigclos$.
+The code memory may contain sets of seals, but only the seals in $\sigrets$ and $\sigcloss$.
 Other than that, code memory only contains instructions in the form of integers.
@@
 
-Define worlds as $\World = \Worldh \times \Worlds \times \Worldfs$.
+Define worlds as $\Wor = \Worldh \times \Worlds \times \Worldfs$.
 They are recursive as they contain world indexed relations.

### 2018-10-08 fully-abstract-comp/popl2019/paper.tex
Address reviewer D comments

@@
 \item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present a new way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
+\item We present a new way to formalize these guarantees based on a novel
+  technique called \textit{fully-abstract overlay semantics} and we prove LSE
+  and WBCF claims. This includes:
   \begin{itemize}
@@
 In this section, we introduce our calling convention \stktokens{} that ensures LSE and WBCF.
-We will gradually explain each of the security measures \stktokens{} take and motivate them with the attacks they prevent.
+We will gradually explain each of the security measures \stktokens{} takes and motivate them with the attacks they prevent.
 
@@
 The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
-However, restricting the stack capability does not help when we in the first call give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
+However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
 Generally speaking, we have no reason to ever trust a stack capability received from a component we do not trust as that stack capability may have been duplicated and stored for later use.
@@
 Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
-The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, the they cannot depend on the same spatially owned region.
+The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, they cannot depend on the same spatially owned region.
 % memory relation

### 2018-10-13 fully-abstract-comp/popl2019/paper.tex
Fix LaTeX compilation.
Add references suggested by Robert Watson...

@@
 
-\usepackage[disable]{todonotes}
+\usepackage{todonotes}
 
@@
 \todo[inline]{Lau: for camera ready version take a look at the coloring in the paper}
-\todo[inlint]{Lau: for camera ready version if TOPLAS paper accepted, update the reference in LR section. }
+\todo[inline]{Lau: for camera ready version if TOPLAS paper accepted, update the reference in LR section. }
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
@@
 \citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
+In fact, Robert Watson told us privately that a similar system was envisioned in earlier work on CHERI~\citep{watson2012cheri}. 
 % KJAA: 'Boundary crossing' is new here.
@@
 
+\todo{Watson et al.\ (XXX: fix name to reflect citation) have concurrently developed an extension of the CHERI-MIPS instruction set with linear capabilities.}
+\todo{\url{https://nakamotoinstitute.org/scarce-objects/}}
+\todo{\url{http://journals.uic.edu/ojs/index.php/fm/article/view/548}}
+
 In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.

### 2018-10-14 fully-abstract-comp/popl2019/paper.tex
[WIP] Updated figure 10 formatting
Still working on the format.

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
   \end{tabular}
-
-  \begin{mathpar}
-    \inferrule{
-      \src{r_1 \neq \rtmp{1}}\\
-      \src{\reg(r_1) = \sealed{\sigma,c_1} }\\
-      \src{r_2 \neq \rtmp{1} }\\
-      \src{\reg(r_2) = \sealed{\sigma',c_2} }\\
-      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} }\\
-      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
-      \src{\ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
-      \src{\ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
-      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
-      \src{c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
-      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) }\\
-      \src{\opc = \aaddr + \calllen }\\
-      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }\\
-      \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }\\
-      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} }\\
-      \src{\baddr \leq  \aaddr+\offpc \leq \eaddr }\\
-      \src{\sigma = \sigma_\aaddr + \offsigma }\\
-      \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }\\
-      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)}}\\ 
-      \src{s_c = \sealed{\sigma,c_\opc} }\\
-      \src{s_d = \sealed{\sigma,c_\var{priv\_data}} }\\
-      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}}
-    {\src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}}
-  \end{mathpar}
 \begin{multline*}
@@
     &\phantom{mak} &  $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
-      & & \\
-    \sourcecolor\left.
-      \arraycolsep=0pt
-      \begin{array}[t]{rl}
-        \Phi' & \updReg{\pcreg}{c_\opc}\\
-              & \updReg{\rstk}{c_\stk}\\
-              & \updReg{\rdata}{0}\\
-              & \updReg{\rtmp{1}}{0}\\
-              & \updReg{\rtmp{2}}{0}
-      \end{array}\right.
+        \src{\Phi'} 
       & &
-        \srcalt{$c_1 = \retptrc(\baddr,\eaddr,\opc)$ and
-        $c_2 = \retptrd(\aaddr_\stk,\eaddr_{\stk})$ and
-        $\Phi(\rstk) = \stkptr{\rw,\stkb,\aaddr_\stk-1,\_}$ and
-        $\stkb \leq \aaddr_\stk-1$ and
-        $\Phi = (\mem,\reg,\stkf::\stk,\ms_\stk)$ and
-        $\stkf = (\opc,\ms_{\stk,\priv})$ and
-        $c_\opc = ((\rx,\normal),\baddr,\eaddr,\opc)$ and
-        $\dom(\ms_{\stk,\priv}) = [\aaddr_\stk,\eaddr_{\stk}]$ and
-        $c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk},\aaddr_\stk}$ and
-        $\Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk)$ }\\
-      & & \\
+        $\src{ \Phi' = \var{returnCondition}(c_1,c_2,\Phi)}$ \\
       \failed &  & \totherwise
@@
   \caption{An excerpt of the operational semantics of \srccm{}.}
+  \label{fig:source-op-sem}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \inferrule{
+      \src{\exists \sigma \ldotp \reg(r_1) = \sealed{\sigma, c_1} \wedge \reg(r_2) = \sealed{\sigma, c_2}} \\
+      \src{r_1 \neq \rtmp{1}} \\ \src{r_2 \neq \rtmp{2}}
+    }{
+      \src{\var{argumentsSealedPairs}(\reg,r_1,r_2,c_1,c_2)}
+    }
+    \and
+    \inferrule{
+      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} }\\
+      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
+      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
+      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
+      \src{\ms_{\stk,\var{unused}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
+      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }
+    }{
+      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk)}
+    }
+    \and
+    \inferrule{
+      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) }\\
+      \src{\opc = \aaddr + \calllen }\\
+      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }
+    }{
+      \src{ \opc, c_\opc = \var{setupOpc(\reg)}}
+    }
+    \and
+    \inferrule{
+      \src{\var{callStackFrame} = (\opc,\ms_{\stk,\priv})}\\
+      \src{\stk' = \var{callStackFrame} :: \stk }
+    }{
+      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}})}      
+    }
+    \and
+    \inferrule{
+      \src{\reg(\pcreg) = ((\_,\_),\_,\_,\aaddr) }\\
+            \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} }\\
+      \src{\baddr \leq  \aaddr+\offpc \leq \eaddr }\\
+      \src{\sigma = \sigma_\aaddr + \offsigma }\\
+      \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }
+    }{
+      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)}
+    }
+    \and
+    \inferrule{
+      \src{s_d = \sealed{\sigma,c_\var{priv\_data}}}\\
+      \src{s_c = \sealed{\sigma,c_\opc}}
+    }{
+     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc)}
+    }    
+    % \and
+    % \inferrule{
+    %   % \src{r_1 \neq \rtmp{1}}\\
+    %   % \src{\reg(r_1) = \sealed{\sigma',c_1} }\\
+    %   % \src{r_2 \neq \rtmp{1} }\\
+    %   % \src{\reg(r_2) = \sealed{\sigma',c_2} }\\
+    %   % \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} }\\
+    %   % \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
+    %   % \src{\ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
+    %   % \src{\ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
+    %   % \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
+    %   % \src{c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
+    %   % \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) }\\
+    %   % \src{\opc = \aaddr + \calllen }\\
+    %   % \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) }\\
+    %   % \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }\\
+    %   % \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} }\\
+    %   % \src{\baddr \leq  \aaddr+\offpc \leq \eaddr }\\
+    %   % \src{\sigma = \sigma_\aaddr + \offsigma }\\
+    %   % \src{\sigma_\baddr \leq \sigma \leq \sigma_\eaddr }\\
+    %   \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk)}\\
+    %   \src{\opc, c_\opc =\var{setupOpc(\reg)}}\\
+    %   \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}})}\\
+    %   \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)}\\
+    %   \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc)}\\
+    %   \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)}}\\ 
+    %   \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
+    % }
+    % {\src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}}
+  \end{mathpar}
+  \[
+    \begin{array}{l}
+    \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})} \quad \mathit{iff}\\
+      \quad\quad\left\{ 
+      \begin{array}{l}
+      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk) \wedge{}}\\
+      \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
+      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}}) \wedge{}}\\
+      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
+      \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc) \wedge{}}\\
+      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
+      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
+      \end{array}\right.
+  \end{array}
+\]
+\[
+    \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
+\]
+\[
+    \src{((\rx,\normal),\baddr,\eaddr,\opc) = \var{tokenToPc}(\retptrc(\baddr,\eaddr,\opc)}
+\]
+\[
+    \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
+    \quad\mathit{iff}\quad
+      \left\{ \array{l}
+        \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
+        \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
+        \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
+        \src{\dom(\ms_\var{local\_stk} \subseteq [\aaddr_\stk,\eaddr_\stk])}
+        \endarray
+     \right.
+\]
+\[
+    \src{c_\stk = \var{reconstructStackPointer}(\reg,c)}
+    \quad\mathit{iff}\quad
+      \left\{ \array{l}
+        \src{c = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge{}} \\
+        \src{\reg(\rstk) = \stkptr{\rw, \stkb, \aaddr_\stk-1,\_} \wedge{}} \\
+        \src{\stkb \leq \addr_\stk \wedge{}} \\
+        \src{c_\stk = \var{tokenToStkPtr(c)} }
+        \endarray
+     \right.
+\]
+  \[
+    \begin{array}{l}
+    \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))} \quad \mathit{iff}\\
+      \quad\quad\left\{ 
+      \begin{array}{l}
+      \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
+      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
+      \src{c_\opc =  \var{tokenToPc}(c_1)} \\
+      \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
+      \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
+      \end{array}\right.
+  \end{array}
+\]
+
+\begin{mathpar}
+  \inferrule{     
+  }{
+d
+  }
+\end{mathpar}
+  \caption{The call condition in the operational semantics of \srccm{}.}
   \label{fig:source-op-sem}

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Update my affiliation.
Update to camera-ready format.
Small grammar fix.

--- 2018-10-13 fully-abstract-comp/popl2019/paper.tex
@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall]{acmart}\settopmatter{}
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
@@
 \affiliation{
-  \institution{KU~Leuven}           %% \institution is required
+  \institution{Vrije~Universiteit~Brussel}           %% \institution is required
   \country{Belgium}                   %% \country is recommended
 }
-\email{dominique.devriese@cs.kuleuven.be}         %% \email is recommended
+\email{dominique.devriese@vub.be}         %% \email is recommended
 
@@
 a piece of code coupled with private state that it gains access to upon invocation.
-The concrete mechanics of object capabilities varies between different capability machines.
+The concrete mechanics of object capabilities vary between different capability machines.
 % KJAA: This 'sealed' and 'common seal' business is unclear to me.

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Some small changes.

@@
 Each of the two capabilities are sealed with a common seal which make them opaque.
-The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015}.
+The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.
 
@@
 For example, CHERI offers a form of {\itshape local} capabilities that can only be stored in registers or on the stack but not in other parts of memory.
-\citet{skorstengaard_reasoning_2017} has demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
-In fact, Robert Watson told us privately that a similar system was envisioned in earlier work on CHERI~\citep{watson2012cheri}. 
+\citet{skorstengaard_reasoning_2017} have demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
+In fact, a similar system was envisioned in earlier work on CHERI~\citep{watson2012cheri}. 
 % KJAA: 'Boundary crossing' is new here.
@@
 We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings in particular without the need for clearing large blocks of memory.
+The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
 
@@
   \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
-  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
+  \item proving full-abstraction for the embedding of \srccm{} into \trgcm{} (Section~\ref{sec:fa-proof}) by
   \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
@@
 % Cap machine inspired by CHERI, but with linear capabilities
+In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
+For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
+% \lau{Do we have a good reason to assume an infinite address space for this work?
+%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
+
+%%% Capabilities
 \begin{wrapfigure}{r}{0.26\linewidth}
@@
 \end{wrapfigure}
-In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
-For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
-% \lau{Do we have a good reason to assume an infinite address space for this work?
-%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
-
-%%% Capabilities
 The concept of a capability is the cornerstone of any capability machine.

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Some small modifications in section 2.

@@
 In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
-If both things are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
+If both conditions are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
 The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
@@
 % load w/ lin (and store)
-The \texttt{store} and \texttt{load} instructions are fairly standard: They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority, and they clear the source of a linear capability transfer.
+The \texttt{store} and \texttt{load} instructions are fairly standard.
+They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority, and they clear the source of a linear capability transfer.
 Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
@@
 % write xor exec
-In Section~\ref{sec:form-secur-with}, we will get back to how a program is instrumented to create an executable configuration.
+In Section~\ref{sec:form-secur-with}, we will get back to how a program is initialised to create an executable configuration.
 % We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Small modifications in section 3.

@@
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
-In both examples our component and some adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
+In both examples our component and an adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
 The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
 However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
-Generally speaking, we have no reason to ever trust a stack capability received from a component we do not trust as that stack capability may have been duplicated and stored for later use.
+Generally speaking, we have no reason to ever trust a stack capability received from an untrusted component as that stack capability may have been duplicated and stored for later use.
 In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
-The adversary may have stored the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed to use it.
+The adversary stores the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed.
 The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.
@@
 % Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
-Because the stack capability is linear the caller knows that the capability for their local stack frame cannot have an alias.
-This means that an adversary would need the stack capability the caller produced
-in order to access the caller's local data.
-While the caller gives the capability to the adversary, it is only after the caller seals it which makes the capability opaque.
+Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
+This means that an adversary would need the stack capability produced by the caller in order to access their local data.
+The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
 This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
@@
 A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
-However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
+However without any enforcement, there is nothing to prevent a callee from returning on an arbitrary call on the call stack.
 This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
@@
 In order for a call to have a presence on the call stack, its stack frame must be non-empty.
-If we could have empty stack frames on the call stack, then it would be impossible to tell whether the topmost non-empty stack frame has an empty stack frame on top of it.
+We cannot allow empty stack frames on the call stack, because then it would be impossible to tell whether the topmost non-empty stack frame has an empty stack frame on top of it.
 Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in \stktokens{} these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
@@
 At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
-However, this does not ensure that the caller's stack frame is on top of the call stack.
+However, this still does not ensure that the caller's stack frame is on top of the call stack.
 The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
-An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they later can return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
+An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
 This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
@@
 For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In \stktokens{}, the base address of the stack capability is fixed as a
-compile-time constant (Note: the stack grows downwards, so the base address of
-the stack capability is the top-most address of the stack). 
-The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
+In \stktokens{}, the base address of the stack capability is fixed as a compile-time constant (Note: the stack grows downwards, so the base address of the stack capability is the top-most address of the stack).
+The caller verifies the validity of the return token by checking whether the base address of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
 In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Small shortenings in section 4.

@@
 % Note about them vs us, we do nothing that they couldn't do
-In the description of the \stktokens{} calling scheme, it has sometimes been in terms of ``them vs us''.
+We have sometimes phrased the description of the \stktokens{} calling scheme in terms of ``them vs us''.
 This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
@@
 % Summary of CC
-In the this section, we have described our proposed calling convention, \stktokens{}. The measures taken are:
+To summarise, our proposed calling convention \stktokens{} consists of the following measures:
 \begin{enumerate}
@@
 \item \label{item:return-data} Create token and data return capability on call: split the stack capability in two to get a stack capability for your local stack frame and a stack capability for the unused part of the stack.
-The stack capability for your local stack frame is sealed and used for the data part of the return pair.
+The former is sealed and used for the data part of the return pair.
 \item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, all return seals are only used for one call site, and return seals are not leaked.
+\item Reasonable use of seals: Return seals are only used to seal old program pointers, all return seals are only used for one call site, and they are not leaked.
 \end{enumerate}

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Small fixes in section 4.

@@
 
-The fact that \stktokens{} enforces LSE and WBCF is formulated as the theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
+The fact that \stktokens{} enforces LSE and WBCF is formulated as a theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
 The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
-Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than the corresponding components in the well-behaved overlay semantics.
+Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than components which the trusted component interacts with in the well-behaved overlay semantics.
 In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
@@
 Such seals should also never be passed to other components.
-Such requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
+These requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
 The general pattern is that $\stktokens{}$ only protects components that do not shoot themselves in the foot by violating a few basic rules.
@@
                \src{\perm \in \{\rwx,\rx\}} \\
-               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2}, \dots, \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
+               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2} \cdots \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
  }
-              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}} }
+              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} }
   \end{mathpar}

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Refactor figure 10

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall]{acmart}\settopmatter{}
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[acmsmall]{acmart}\settopmatter{}
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
   \label{fig:source-op-sem}
+\end{figure}
+
+\begin{figure}
+    \[
+      \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)} \mathit{ iff }
+      \left\{ 
+      \begin{array}{l}
+        \src{\exists \sigma \ldotp (\reg(r_1) = \sealed{\sigma, c_1} \wedge \reg(r_2) = \sealed{\sigma, c_2}) \wedge{}}\\
+        \src{r_1 \neq \rtmp{1} \wedge r_2 \neq \rtmp{2}}
+      \end{array}
+      \right.
+    \]
+    \[
+      \src{ \opc, c_\opc = \var{setupOpc(\reg)}} \mathit{ iff }
+      \left\{ 
+      \begin{array}{l}
+      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge{}}\\
+      \src{\opc = \aaddr + \calllen \wedge{}}\\
+      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
+      \end{array}
+       \right.
+    \]
+    \begin{multline*}
+      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk)} \mathit{\ iff} \\
+      \left\{ 
+      \begin{array}{l}
+      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \wedge \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
+      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
+      \src{\ms_{\stk,\var{unused}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
+      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }
+      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
+      \end{array}
+      \right.
+    \end{multline*}
+    \[
+      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}})}\text{\ iff\ }
+            \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }
+    \]
+    \[
+      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)} \mathit{\ iff\ }
+      \left\{
+      \begin{array}{l} 
+      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
+      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge{} \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \wedge{}}\\
+      \src{\sigma = \sigma_\aaddr + \offsigma }
+      \end{array}
+      \right.
+    \]
+
+    \[
+     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc)} \mathit{\ iff\ }
+      \src{s_d = \sealed{\sigma,c_\var{priv\_data}} \wedge{} s_c = \sealed{\sigma,c_\opc}}
+    \]
+  \begin{multline*}
+    \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}\mathit{\ iff}\\
+      \quad\quad\left\{ 
+        \begin{array}{l}
+          \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)}\\
+      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk) \wedge{}}\\
+      \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
+      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}}) \wedge{}}\\
+      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
+      \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc) \wedge{}}\\
+      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
+      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
+      \end{array}\right.
+  \end{multline*}
+
+  \caption{The call condition in the operational semantics of \srccm{}.}
+  \label{fig:source-op-sem}
+\end{figure}
+\begin{figure}
+\[
+    \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
+\]
+\[
+    \src{((\rx,\normal),\baddr,\eaddr,\opc) = \var{tokenToPc}(\retptrc(\baddr,\eaddr,\opc)}
+\]
+\[
+    \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
+    \mathit{\ iff\ }
+      \left\{ \array{l}
+        \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
+        \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
+        \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
+        \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
+        \endarray
+     \right.
+\]
+\[
+    \src{c_\stk = \var{reconstructStackPointer}(\reg,c)}
+    \mathit{\ iff\ }
+      \left\{ \array{l}
+        \src{c = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge{}} \\
+        \src{\reg(\rstk) = \stkptr{\rw, \stkb, \aaddr_\stk-1,\_} \wedge{}} \\
+        \src{\stkb \leq \addr_\stk \wedge{}} \\
+        \src{c_\stk = \var{tokenToStkPtr(c)} }
+        \endarray
+     \right.
+\]
+
+\begin{multline*}
+    \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
+      \quad\quad\left\{ 
+      \begin{array}{l}
+      \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
+      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
+      \src{c_\opc =  \var{tokenToPc}(c_1)} \\
+      \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
+      \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
+      \end{array}\right.
+  \end{multline*}
+
+  \caption{The return condition in the operational semantics of \srccm{}.}
+  \label{fig:source-op-sem-return}
 \end{figure}

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Add inline comments to code in figure 8.

@@
 \[
-  \begin{array}{r >{$}p{3.5cm}<{$} l >{$}p{3.5cm}<{$} l l}
-    1: & \tmove{\rtmp{1}}{42} & 10: &\tcseal{\rretd}{\rtmp{1}} & 19: &\tcca{\rtmp{2}}{5} \\
-    2: &\tstore{\rstk}{\rtmp{1}} & 11: &\tmove{\rretc}{\pcreg} & 20: &\tjnz{\rtmp{2}}{\rtmp{1}}\\
-    3: &\tcca{\rstk}{(-1)} & 12: & \tcca{\rretc}{5} & 21: &\tcca{\rtmp{2}}{1} \\
-    4: &\tgeta{\rtmp{1}}{\rstk} & 13: & \tcseal{\rretc}{\rtmp{1}} & 22: &   \tjmp{\rtmp{2}} \\
-    5: &\tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}} & 14: & \tmove{\rtmp{1}}{0} & 23: &    \tfail \\
-    6: &\tmove{\rtmp{1}}{\pcreg} & 15: & \txjmp{r_1}{r_2} & 24: &   \tsplice{\rstk}{\rstk}{\rdata} \\
-    7: &\tcca{\rtmp{1}}{(\offpc - 5)} & 16: & \tgetb{\rtmp{1}}{\rstk} & 25: &    \tcca{\rstk}{1}\\
-    8: &\tload{\rtmp{1}}{\rtmp{1}} & 17: &\tminus{\rtmp{1}}{\rtmp{1}}{\stkb} & 26: &    \tmove{\rtmp{2}}{0} \\
-    9: &\tcca{\rtmp{1}}{\offsigma} & 18: &\tmove{\rtmp{2}}{\pcreg} & &
+  \begin{array}{r >{$}p{5.1cm}<{$} r l}
+\multicolumn{2}{l}{\text{// Ensure non-empty stack.}}            & \multicolumn{2}{l}{\text{// Clear tmp registers and jump.}}\\
+1 :& \tmove{\rtmp{1}}{42}                                        &14: & \tmove{\rtmp{1}}{0}\\
+2 :& \tstore{\rstk}{\rtmp{1}}                                    &15: & \txjmp{r_1}{r_2}\\
+3 :& \tcca{\rstk}{(-1)}                                          &\multicolumn{2}{l}{\text{// The following is the return code.}}\\
+\multicolumn{2}{l}{\text{// Split stack in local stack frame and unused.}}      &\multicolumn{2}{l}{\text{// Check that returned stack pointer has base $\stkb$.}}\\
+4 :& \tgeta{\rtmp{1}}{\rstk}                                     &16: & \tgetb{\rtmp{1}}{\rstk}\\
+5 :& \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}                     &17: & \tminus{\rtmp{1}}{\rtmp{1}}{\stkb}\\
+\multicolumn{2}{l}{\text{// Load the call seal.}}                &18: & \tmove{\rtmp{2}}{\pcreg}\\
+6 :& \tmove{\rtmp{1}}{\pcreg}                                    &19: & \tcca{\rtmp{2}}{5}\\
+7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
+8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \rtmp{2}{1}\\
+9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
+\multicolumn{2}{l}{\text{// Seal the local stack frame.}}        &23: & \tfail \\
+10:& \tcseal{\rretd}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Splice with capability for local stack frame.}}\\
+\multicolumn{2}{l}{\text{// Construct code return pointer.}}     &24: & \tsplice{\rstk}{\rstk}{\rdata} \\
+11:& \tmove{\rretc}{\pcreg}                                      &\multicolumn{2}{l}{\text{// Pop 42 from the stack}}\\  
+12:& \tcca{\rretc}{5}                                            &25: & \tcca{\rstk}{1}\\
+13:& \tcseal{\rretc}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Clear tmp register}}\\
+      &    &26: & \tmove{\rtmp{2}}{0}\\
   \end{array}

### 2018-10-15 fully-abstract-comp/popl2019/paper.tex
Revert figure 8 caption.

@@
 \caption{
-  The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
-  Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
-  Instructions 1-3: ensure that the stack is non-empty.
-  Instructions 4-5: split stack pointer in two (stack pointer for callee done).
-  Instructions 6-9: Load the seal designated for this call.
-  Instruction 10: Seal capability for private stack to create data return pointer.
-  Instructions 11-13: Seal capability for return point to create code return pointer.
-  Instructions 14-15: Clean up temporary register and jump.
-  Instructions 16-26: The return code.
-  Instructions 16-23: check the stack base.
-  Instruction 24: join the stack pointer.
-  Instructions 25-26: pop 42 from the stack.
-%The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
-The variable $\offsigma$ is the offset in the set of seals.
+  % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
+  % Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
+  % Instructions 1-3: ensure that the stack is non-empty.
+  % Instructions 4-5: split stack pointer in two (stack pointer for callee done).
+  % Instructions 6-9: Load the seal designated for this call.
+  % Instruction 10: Seal capability for private stack to create data return pointer.
+  % Instructions 11-13: Seal capability for return point to create code return pointer.
+  % Instructions 14-15: Clean up temporary register and jump.
+  % Instructions 16-26: The return code.
+  % Instructions 16-23: check the stack base.
+  % Instruction 24: join the stack pointer.
+  % Instructions 25-26: pop 42 from the stack.
+  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$.
+  The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
+  The variable $\offsigma$ is the offset in the set of seals to the call seal.
   The constant $\stkb$ is the globally agreed on stack base.
   There are a number of ``magic'' numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
-Line 7: $-5$, offset to the first instruction (from the point the capability was moved out of the $\pcreg$-register).
+  Line 7: $-5$, offset to the first instruction of the call (from the point the capability was moved out of the $\pcreg$-register).
   Line 12: $5$, offset to the return address.

### 2018-10-16 fully-abstract-comp/popl2019/paper.tex
Rewrite and simplify Section 5
Specific simplifications:
* do not define the code region
* remove all mention of step-indexing or approximation direction
* other minor simplifications wherever possible.

--- 2018-10-15 fully-abstract-comp/popl2019/paper.tex
@@
 % Paper specific redefinitions of commands
+\renewcommand{\npair}[2][n]{#2}
 \renewcommand{\MemSeg}{\shareddom{MemSeg}}
@@
 \renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
+
+\renewcommand{\lre}[1][]{\typesetlr{E}^{#1}}
+\renewcommand{\lretrusted}[1][]{\lreg[#1]{\trusted}}
+\renewcommand{\lrexj}[1][]{\typesetlr{E}^{#1}_{\mathrm{xjmp}}}
+\renewcommand{\lrk}[1][]{\typesetlr{K}^{#1}}
+\renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
+\renewcommand{\lrr}[1][]{\lrrg[#1]{\untrusted}}
+\renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
+\renewcommand{\lro}[1][]{\typesetlr{O}^{#1}}
+\renewcommand{\lrol}{\lro[\preceq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
+\renewcommand{\lror}{\lro[\succeq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
+\renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
+\renewcommand{\lrv}[1][]{\lrvg[#1]{\untrusted}}
+\renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
+\renewcommand{\lrp}{\typesetlr{P}}
+
+\renewcommand{\lrrs}{\typesetlr{R}}
+\renewcommand{\lrm}{\typesetlr{M}}
+\renewcommand{\lrcomp}[1][]{\typesetlr{C}^{#1}}
+\renewcommand{\lrec}[1][]{\typesetlr{EC}^{#1}}
+\renewcommand{\lrheap}[1][]{\typesetlr{H}^{#1}}
+
+\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{\gc}}
+\renewcommand{\stdreg}[3][]{\iota^{\mathrm{std},#3,#1}_{#2}}
+\renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
+\renewcommand{\codereg}[2][\mathrm{code}]{\iota^{#1}_{#2}}
+\renewcommand{\staureg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta,\lrv},#2}_{#1}}
 
@@
 \subsection{Well-formedness and reasonability}
+\label{sec:well-form-reas}
 % Well-formed components
@@
 % \end{itemize}
-The backbone of the full-abstraction proof is a cross-language, step-indexed, Kripke logical relation.
-It relates the observable behavior of \srccm{} to the observable behavior of \trgcm{}.
-In the full-abstraction proof, it allows us to transfer termination assumptions between the two machines which enables the use of a contextual equivalence assumption.
-We first give an informal presentation of the logical relation before we return to a sketch of the full-abstraction proof.
-
-\subsection{Logical relation}
-\label{subsec:logical-relation}
-In this section, we present the logical relation.
-Unfortunately, for space reasons, we do not give a comprehensive description of it.
-Instead we highlight specific parts to the logical relation and refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
-Finally, for presentation, we will sometimes omit details, like step indexing, that are important for correctness but otherwise not interesting.
-
-\subsubsection{Worlds}
+To prove Theorem~\ref{thm:full-abstraction}, we will essentially show that trusted components in \srccm{} are related in a certain way to their embeddings in \trgcm{}, and that untrusted \trgcm{} components are similarly related to their embeddings in \srccm{}.
+We will then prove that these relations imply that the combined programs have the same observable behavior, i.e.\ one terminates iff the other does.
+The hard part is in defining when components are related.
+In the next section, we give an informal overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
+
+\subsection{Kripke Worlds}
 \label{subsec:worlds}
-Generally speaking, a world is used in a logical relation to model assumptions about memory.
-\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
-All of the sub-worlds are partial maps from $\RegionName$ (modelled as natural numbers) to regions.
-The regions differ for each of the sub-worlds based on what they model.
-Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
-The region types model linearity as they indicate how the memory can be referenced.
-We have four types of regions: $\pure$, $\spatial$, $\spatialo$, and $\revoked$.
-The $\pure$ regions can be referenced by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ and $\revoked$ regions by nothing.
-To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
-To this end, we have the partial operation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except that they have disjoint ownership: $\pure$ regions can be present in both worlds, but a $\spatialo$ region can only be present in one.
-Such a region must then be $\spatial$ in the other world.
-%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
-This is important because the logical relation will require that valid memories are backed by a suitable memory segment even when the memory does not own them (for example, when a linear capability for it is in one of the registers).
-Revocation of linear capabilities is supported in a certain way using the $\revoked$ region.
-Finally, $\pure$ regions can also claim ownership of a seal by defining a seal interpretation for it.
-The seal interpretation is a world-indexed relation that must hold for all sealables signed by this seal.
-We split the regions in two categories:
-\begin{align*}
-  \Regionh &= 
-             \left\{\begin{multlined}
-               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
-               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
-             \end{multlined}\right.\\
-  \Regions &={}
-             \left\{\begin{multlined}
-               \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
-               \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
-               \{\revoked\}
-             \end{multlined}\right.
-\end{align*}
-and define the sub-worlds as follows:
+The relation between \srccm{} and \trgcm{} components is non-trivial: essentially, we will say that components are related if invoking them with related values produces related observable behavior.
+However, values are often only related under certain assumptions about the rest of the system.
+For example, the linear data part of a return capability should only be related to the corresponding \srccm{} capability if no other value in the system references the same inactive stack frame and if it is sealed with a seal that is only used for return pointers to the same code location.
+To accomodate such conditional relatedness, we construct our relation as a step-indexed Kripke logical relation with recursive worlds.
+Space constraints prevent us from explaining it in full detail in this paper. 
+Instead, we will only highlight specific important parts of the logical relation in this section, and we refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
+Additionally, for presentation, we will omit details, like step indexing, that are important for correctness but otherwise uninteresting.
+
+Assumptions about the system, that relatedness is predicated on, are gathered in (Kripke) worlds. 
+We use a type of worlds that is tailored specifically for our purposes.
+They consist of three sub-worlds: $\Wor = \Worldh \times \Worlds \times \Worldfs$, capturing assumptions about the heap\footnote{Actually, we use the term heap to describe all memory except the stack, including, for example, code memory.}, the inactive and the active part of the stack, respectively.
+Sub-worlds consist of a finite mapping from region names to regions, which come in two forms (spatial and shared):
 \begin{align*}
@@
 \end{align*}
-The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
-The call stack and free stack are only ever addressed by linear capabilities, so their sub-worlds only allow $\Regions$.
-Regions in $\Worlds$ correspond to a stack frame, so they contain the return address for this frame.
-
-Define worlds as $\Wor = \Worldh \times \Worlds \times \Worldfs$.
-They are recursive as they contain world indexed relations.
-Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct $\Wor$ which solves the circularity\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
+Different parts of the world can contain different types of regions: heap-related assumptions can be either shared or spatial (see below), while stack-related regions must be spatial.
+Additionally, regions for inactive parts of the stack additionally include an address specifying the return address for that stack frame.
+
+Regions in $\Regionh$ specify the presence of an invariant in the system, shared with the rest of the system.
+\begin{align*}
+  \Regionh &= 
+             \left\{\begin{multlined}
+               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg}) \times \\
+               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
+             \end{multlined}\right.\\
+  \Regions &={}
+             \left\{\begin{multlined}
+               \{\spatial,\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \{\revoked\}
+             \end{multlined}\right.
+\end{align*}
+Shared regions are tagged with the syntactic token $\pure$ and may prescribe two different types of requirements.
+First, they may require the presence of \srccm{} and \trgcm{} memory segments satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$ (see below for the meaning of $\Wor \monnefun$).
+A region might, for example, require the presence of a certain list of instructions at a given set of memory addresses in both \srccm{} and \trgcm{}.
+If a memory segment is owned by a given region, then that memory segment must be disjoint from the memory segments owned by all other regions.
+Second, a shared region may also contains a partial function from seals to relations on pairs of sealable capabilities: $\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps}$ (see below for the meaning of $\Wor \monnefun$).
+When the region defines such a relation for a given seal, then no other region in the world can do the same, and any value signed with that seal will be required to satisfy the registered relation.
+
+Spatial regions are similar to shared regions, but they are tagged as $\spatial$ or $\spatialo$ and may not specify seal invariants.
+Additionally, a spatial region may also be $\revoked$.
+The difference between spatial and shared regions is related to linearity and ownership.
+For example, a \trgcm{} linear capability to a piece of memory is related to its \srccm{} counterpart, but only if no other linear capability overlaps with it.
+We will model such an assumption for exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region, that only one value in the system may rely on.
+More concretely, we will define how to combine worlds $W_1$ and $W_2$ into a combined world $W_1 \oplus W_2$, on the condition that they represent compatible assumptions: $W_1$ and $W_2$ must contain the same regions except that they must respect exclusive ownership of $\spatialo$ regions: $\pure$ regions must be present in both worlds, but a $\spatialo$ region can only be present in one and must be $\spatial$ in the other.
+We say that worlds $W_1$ and $W_2$ are compatible when $W_1 \oplus W_2$ is defined.
+
+The reason that we have $\spatial$ regions is for defining when \srccm{} and \trgcm{} memories are related.
+We need them to contain suitable memory segments for all regions in the system, even for regions owned by values that live outside the memory (for example register values).
+Those regions will be in the world for the memory, but only as $\spatial$, i.e.\ they may not actually be referenced from within memory, but we still require them to be backed by suitable memory contents.
+We also use a relation $W_2 \future W_1$ that defines when a world $W_2$ is a future world of a world $W_1$.
+Relations that hold with respect to $W_1$, will then generally continue to hold in $W_2$.
+Our future world relation is fairly standard: the future world must contain all the previous world's regions, except that $\spatial$ regions are allowed to become $\spatialo$ (i.e.\ gaining ownership of a region will never break relatedness) or $\revoked$ (i.e.\ revoking spatial regions will never break relatedness).
+
+Attentive readers may have noticed that our definition of worlds is actually cyclic: worlds in $\Wor$ contain regions in $Regionh$, but those contain partial functions from the set $\Wor$ to something else.
+Such recursive worlds are in fact common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} (essentially an advanced form of step-indexing) to construct the set $\Wor$ and rigorously resolve the circularity.
 % \begin{lemma}
@@
 % \end{lemma}
-The future world relation $\future$ is fairly standard.
-It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
-The $\spatial$ regions can become $\spatialo$ or $\revoked$.
-
-\subsubsection{Code region}
-The code region is a central region to the logical relation that models the code memory of a component.
-It is especially interesting because its seal interpretation function specifies what the return and closure seals in a component can seal.
-The code region is defined as follows:
-\[
-  \codereg{\sigrets,\sigcloss,\mscode,\gc} \defeq (\pure,
-  H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \mscode \; \gc,
-  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)
-\]
-The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
-The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself, i.e.\ our component's code memory must not be modified.
-$H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
-As we will see later, we have two value relations: a trusted and an untrusted.
-The value relation used depends on whether the code region governs trusted or untrusted code.
-For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation additionally includes things that trusted components are allowed to have but not pass away, e.g.\ return seals.
-
-% \todo[inline]{Describe $H^\mathrm{code}$}
-% \begin{multline*}
-%   H^\mathrm{code} \; \sigrets \; \sigcloss \; \mscode \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
-%   \left\{
-%     \begin{array}{l}
-%       \npair{(\mscode \uplus \mspad, \mscode \uplus \mspad)} \mid
-%       \dom(\mscode) = [\baddr,\eaddr] \tand  \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \tand\\
-%       ([\baddr - 1, \eaddr + 1] \subseteq \ta \tand
-%       \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \tor \\
-%       \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \tand \\
-% %      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
-%       \forall a \in \dom(\mscode)\ldotp \npair{(\mscode(a),\mscode(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
-%     \end{array}
-%   \right\}
-% \end{multline*}
-
-The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigcloss$ can seal.
-The return seals can only be used to seal return pointers that corresponds to the return from a call in the code of the region.
-Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
-The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (Figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
-The code return pointer corresponds to a normal capability which means that it can be reused but only to return from the same call.
-Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
-Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function relates data return pointers to linear capabilities.
-It also requires the data return pointer to refer to a $\spatialo$ region for one of the stack frames on the call stack.
-The return address of this region in $\Worlds$ must be the return address of the call that the return seal is affiliated with.
-\begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \; \sigma \; \hat{W}= \\
-  \begin{array}[t]{l}
-\left\{
-    \begin{array}{l}
-      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\mathrm{xjmp\_off}+1)\right)} \mid \dots
-    %     \sigrets \subseteq \gsigrets \tand \\
-    %     \quad\dom(\mscode) \subseteq \ta \tand 
-    %     \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-    %     \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
-    %     \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
-    %     \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
-    \end{array}
-      \right\} \cup \\
-\left\{
-    \begin{array}{l}
-      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid \dots
-        % \sigrets \subseteq \gsigrets \tand 
-        % \dom(\mscode) \subseteq \ta \tand\\
-        % \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
-        % \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
-        % \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
-        % \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
-        % \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
-    \end{array}
-    \right\} \\
-  \end{array}\\
-    \text{ if } \sigma \in \sigrets
-\end{multline*}
-
-Unlike return seals, closure seals are allowed in both trusted and untrusted components, but we cannot allow them to use closure seals to seal the same things.
-If we allowed this, then a trusted component could seal a read-only capability for its code (i.e.\ access to return seals) with an untrusted closure seal and give it to the untrusted component which can unseal it with \texttt{xjmp}.
-The closure seals in untrusted code should therefore only be used to seal things from the untrusted value relation, i.e.\ things that are safe to transfer to an untrusted component.
-The trusted closure seals can be used to seal things from the untrusted value relation.
-This allows trusted seals to seal capabilities for the trusted code which is necessary in order to create a new closure.
-
-% and
-% \begin{multline*}
-%   H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \;
-%   (\ta,\stkb) \; \sigma \; \hat{W}= \\
-%   \left\{
-%     \begin{array}{l}
-%  \npair{(\vsc, \vsc' )} \mid
-%         (\dom(\mscode) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
-%         (\dom(\mscode) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
-%          \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
-%          \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
-%     \end{array}
-%   \right\}\\
-%   \text{ if } \sigma \in \sigcloss
-% \end{multline*}
-
-\subsubsection{Expression relations and friends}
-The logical relation has two expression relations.
-They both relate capabilities that can safely be used for execution in configurations that satisfy a given world.
-The first one, $\mathcal{E}$, is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
-In our case, the only observable thing on the capability machines is termination.
-The observation relation $\mathcal{O}$ (Figure~\ref{fig:obs-rel}) defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations.
-However, the only difference between the two approximations is in the observation approximation.}.
-\begin{figure}[htb]
-  \centering
-  \begin{align*}
-  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
-  \forall i \leq n \ldotp
-    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
-  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
-  \forall i \leq n \ldotp 
-    \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
-\end{align*}
-\caption{The observation approximations.}
-\label{fig:obs-rel}
-\end{figure}
-With sealed capabilities in the mix, we also need to be able to say that \texttt{xjmp}s are safe, but a sealed code capability only needs to be safe to execute when it has access to the related data capability.
-Further, \texttt{xjmp}s are used to invoke return pairs which on \srccm{} is somewhat involved.
-This is handled uniformly by the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$.
-It relates two pairs of capabilities by plugging them into related register-files and memories using the $\xjmpres{}$ function and requiring that the result is in the observation relation.
-\extend{Write that it is untrusted register files + intuition for this}
-  % \begin{multline*}
-  %   \lre(W)= \\
-  %   \left\{ \begin{array}{l}
-  %     \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} | \\
-  %     \quad\forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-  %     \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
-  %     \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}} \tand\\
-  %     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}} \tand W \oplus W_\lrrs \oplus W_\lrm\\
-  %     \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-  %   \end{array}
-  %   \right\}
-  % \end{multline*}
-% \begin{multline*}
-%   \lrexj(W) =
-%   \\\left\{\begin{array}{l}
-%       \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
-%       \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
-%       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-%       \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
-%       \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-%     \end{array}
-%     \right\}
-% \end{multline*}
-% register file relation
-
-Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
-The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, they cannot depend on the same spatially owned region.
-% memory relation
-The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
-This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that correspond with the view \srccm{} provides on the memory.
-Each of the three partitions of $\ms_T$ must be related to one of the three parts of \srccm{} memory under the relevant sub-world of $W$.
-The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
-The free stack $\ms_\stk$ also relates to the free stack partition of $\ms_T$ in the standard way but with respect to the free stack world.
-%Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
-The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts related to each of the stack frames.
-Further, as each of the stack frames corresponds to a slot in the return order none of the stack frames should be empty.
-It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved by the partitions of $\ms_T$.
-
-% %the topmost stack frame should have memory addresses smaller than (the stack grows downwards) the addresses all of the other stack frame, the next and so on.
-% A memory on \trgcm{} $\ms_T$ is related to a heap, call stack and free stack on \srccm{} if the following is possible: $\ms_T$ can be split in three disjoint parts.
-% One part relates to $\ms_S$ with respect to the heap world in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
-% The second part should relate to the free stack $\ms_\stk$ also in the standard way, but with respect to the free stack world.
-% The final part relates to the call stack $\stk$ which means that it can be further split into parts that each are related to one of the stack frames of $\stk$ with respect to the call stack world.
-% The partition must be done in a way that preserves the order of the stack frames in $\stk$.
-% % , so the top stack frame relates to a part of memory where all the other parts are above in the address space, the second stack frame relates to a part of memory where all the other parts are above in the address space with the exception of the topmost part, and so on.
-% % Further, all stack frames should be in addresses above $\stkb$.
-% The seal interpretation functions are in the pure regions, so the heap relation also makes sure that only one seal interpretation function has an interpretation for each seal.
-% \extend{Write about the particulars of the three memory satisfaction: heap relation makes sure only one interpretation for each seal. Free stack $\stkb$}
-% The \texttt{xjmp} expression relation requires register files and memories related under different worlds, namely $W_\lrrs$ and $W_\lrm$, that are related with $\oplus$.
-% This means that the two worlds have the same regions, but distribute the ownership of partial regions between them.
-% By doing so, they make sure that for any part of memory only one of them can have a linear capability it.
-% The two pairs of capabilities related by $\lrexj$ are plugged into their respective configurations using the $\xjmpres{}$ function, so it corresponds to how an $\texttt{xjmp}$ would insert them.
-
-\subsubsection{Value relations}
-The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
-As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonably, e.g.\ by handling return seals in a safe way.
-This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and want to guarantee LSE and WBCF against.
-This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and an untrusted value relation.
-The two value relations are sketched in Figure~\ref{fig:value-relation}.
-
-The untrusted value relation relates words from the two machines that can safely be given to untrusted code.
-Data is always safe, and a piece of data relates to itself.
-A stack pointer is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
-A safe stack pointer with read permission should only be able to produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a safe stack pointer with write permission can at least write safe words.
-Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack sub-world.
-Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read (and write) permission.
-The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
-An executable memory capability should always point to code which means that there should be a code region governing the memory it references.
-The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seals is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
-% The write condition is called $\stackWriteCond{}$
-% \begin{multline*}
-%   \stackWriteCond{W} =\\
-%   \left\{
-%     \begin{array}{l}\npair{A} \mid 
-%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
-%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
-%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
-%     \end{array}
-%   \right\}
-% \end{multline*}
-When a piece of untrusted code is called, it is passed a return pair consisting of two sealed capabilities.
-Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
-The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\src{\vsc_S},\vsc_T)$ must be in the seal interpretation function for $\sigma$.
-Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.
-
-\begin{figure}[htb]
+
+In our proof, we use only a few regions.
+For space reasons, we do not go into their definitions (see the \citet{technical_report}), but we give a brief overview here.
+First, we have the code region $\codereg{\sigrets,\sigcloss,\mscode}$.
+This is a shared region that represents the assumption that memory segment $\mscode$ is loaded in heap memory at a certain location, that it is well-formed and uses return seals $\sigrets$ and closure seals $\sigcloss$.
+It takes ownership of those seals and registers appropriate invariants on the capabilities that may be signed with them.
+
+A second type of regions $\stdreg{A}{\pure}$ and $\stdreg{A}{\spatialo}$ governs heap or stack memory at a set of addresses $A$.
+It simply requires presence of a memory segment for those addresses, such that the memory contains values that themselves satisfy the relation between values that we will see below.
+The contents of the memory is allowed to change as long as the new contents are still valid.
+Finally, a third type of regions $\stareg[\stpair{\ms}{\ms}]{\pure}$ or $\stareg[\stpair{\ms}{\ms}]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
+This region is used, for example, to govern inactive parts of the stack whose contents is required to remain unmodified.
+
+\subsection{The logical relation}
+Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
+The most important relations are summarised in the following table:\\
+% latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
+\newcolumntype{R}{>{$}r<{$}}
+\newcolumntype{L}{>{$}l<{$}}
+\newcolumntype{M}{R@{${}\in{}$}L}
+\begin{tabular}{|M|c|c|}
+  \hline
+  \multicolumn{2}{|c|}{General form} & Relates ... & and ...\\
+  \hline
+  \npair{(w_S,w_T)} & \lrv(W) & values & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\
+  \npair{(w_S,w_T)} & \lrvtrusted(W) & values & \\
+  \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\
+  \npair{(\reg_S,\reg_T)}  &  \lrrtrusted(W) & register files & \\
+  \npair{\Phi_S,\Phi_T}  &  \lro & execution configurations & \\
+  \npair{(w_S,w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
+  \npair{((w_{S,1},w_{S,2}),(w_{T,1},w_{T,2}))}  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
+  \multicolumn{2}{|c|}{$\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$} & memory & \parbox{3.3cm}{satisfy the\\ assumptions in $W$}\\
+  \hline
+\end{tabular}\\
+These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another (omitted) use of step-indexing.
+For space reasons, we cannot show all of these definitions, but we will try to give a good overview, omitting many details.
+
+Note first how we have two value relations, whose definitions are shown in Figure~\ref{fig:value-relation}.
+The difference is that the untrusted value relation $\lrv(W)$ does not just express that the two values are related, but also that they are safe to pass to an untrusted adversary, i.e.
+they cannot be used to break LSE and WBCF.
+The trusted value relation does not have the latter requirement and is a strict superset of the former.
+
+Both relations trivially includes numbers $(i,i)$ which are always related to themselves.
+The untrusted value relation also includes stack pointers and the underlying linear capability (with the same (non-executable) permission, range of authority, and current address), as well as syntactically equal memory capabilities, seals and sealed values, all under certain conditions involving the world $W$ and the capability's properties.
+
+Details are in the \cite{technical_report}, but roughly, for stack capabilities, the omitted condition requires that the world contains a $\spatialo$ region governing this part of the stack.
+For memory capabilities $((\perm,\lin),\baddr,\eaddr,\aaddr)$, a region in the world must govern memory $[\baddr,\eaddr]$, either $\spatialo$ or $\pure$, depending on the linearity $\lin$ of the capability.
+If the capability is executable ($\perm \in \{\rx,\rwx\}$), then we additionally require that the governing region is a code region and that the two capabilities are related $\mathrm{jmp}$ targets, as expressed by the relation $\lre(W)$, in any future world (see below).
+
+Seals allocated to trusted code are related to themselves only by $\lrvtrusted(W)$, but other seals are in both value relations.
+Sealed values are in both relations essentially when the sealed values satisfy the relation that was registered for the seal in a region of the world.
+Additionally, when they are combined with any other pair of values related by that relation, the sealed values are required to be related as $\txjmp{}{}$ targets (i.e. in $\lrexj(W)$).
+Finally, capabilities to code memory are related to themselves in the trusted value relation ($\lrvtrusted(W)$) when there is an appropriate code region in the world.
+They are not in the untrusted value relation because the code memory contains copies of the return seals used by the code, which must not end up in the hands of an adversary.
+
+\begin{figure}
   \centering
@@
       \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} \mid 
-      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss) \wedge \dots 
+      % [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss) \wedge 
+                       \dots 
       % \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
@@
       \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} \mid
-      [\sigma_\baddr,\sigma_\eaddr] \subseteq(\sigrets \cup \sigcloss) \wedge \dots 
+      % [\sigma_\baddr,\sigma_\eaddr] \subseteq(\sigrets \cup \sigcloss) \wedge 
+      \dots 
     %           \exists r \in \dom(\pwheap) \ldotp \\
@@
 
-The trusted value relation includes everything that the untrusted value relation contains and a few extra things.
-Trusted code uses the return seals and the closure seals (in a responsible way), and therefore sets of seals with trusted return seals and closure seals are in the trusted value relation.
-In order for trusted code to even execute, it needs to use an executable memory capability for the trusted code, so such capabilities are in the trusted value relation.\footnote{Note that code capabilities for trusted code is not allowed in the untrusted value relation as it could be used to read return seals that are not in the untrusted value relation.}
-
-% \[
-%   \xReadCond{W} = \left\{ \npair{A} \middle|
-%     \begin{array}{l}
-%       \exists r \in \addressable{\normal, \pwheap} \ldotp \\
-%        \pwheap(r) \nequal \codereg{\_,\_,\mscode,\gc}\tand
-%        \dom(\mscode) \supseteq A 
-%     \end{array}
-%   \right\}
-% \]
-
-\extend{Write some intuition about monotonicity, owned spatial regions and revoked regions}
-
-\subsubsection{Fundamental theorem}
-With the logical relation defined, we state the fundamental theorem of logical relations (FTLR).
-\begin{lemma}[FTLR]
+\begin{figure}
+  \centering
+\begin{align*}
+  \lrrg{\trust}(W) &= \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+    \begin{array}{l}
+      \exists S : (\RegName \setminus \{\pcreg \})\fun \World \ldotp \\
+      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg \} \cup R))} S(r) \wedge \\
+      \quad \forall r \in \RegName \setminus \{\pcreg \}\ldotp \npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
+    \end{array}
+            \right\}\\
+  \lre(W)&= \left\{ \begin{array}{l}
+    \npair{\stpair[.]{w_{c,S}}{w_{c,T}}} | \\
+    \quad\forall \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+    \qquad\npair{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
+    \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{w_{c,S}} \tand\\
+                     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{w_{c,T}} \tand \\
+                     \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+    \qquad\qquad\Rightarrow\npair{\left(\Phi_S', \Phi_T' \right)}\in \lro
+  \end{array}
+  \right\}
+\end{align*}
+  \begin{align*}
+  \lro[] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+    \src{\Phi_S \sterm{}} \Leftrightarrow \Phi_T \trg{\term} \}
+\end{align*}
+\caption{Simplified sketches of the register file relation $\lrr(W)$, the relation
+  for $\com{jmp}$ targets $\lre(W)$  and the observation relation $\lro(W)$.}
+\label{fig:obs-rel}
+\end{figure}
+
+In Figure~\ref{fig:obs-rel}, we show sketches of the register file relation $\lrr$, the relation between $\tjmp{}$ targets $\lre(W)$ and the observation relation $\lro$.
+The trusted/untrusted register file relation simply requires that all registers except $\pcreg$ are in the corresponding value relation (in compatible worlds).
+Two execution configurations are in the observation relation $\lro$ if one terminates whenever the other does\footnote{The actual definition in the \cite{technical_report} is complicated a bit by step-indexing and the fact that we actually use two separate observation relations for left- and right-approximation.}.
+The $\lre(W)$ relation then includes any two words which can be plugged into related register files and memories (in compatible worlds), to obtain execution configurations in the observation relation.
+
+\subsection{Fundamental theorem}
+An important lemma in our proof of full abstraction of the embedding of \srccm{} into \trgcm{}, is the fundamental theorem of logical relations (FTLR).
+The name indicates that it is an instance of a general pattern in logical relations proofs, but is otherwise unimportant.
+\begin{lemma}[FTLR (roughly)]
   \label{thm:ftlr}
-  If\ $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
-  and either $[\baddr,\eaddr] \subseteq \ta$ and\\ $\src{((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
+  If $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and $\pwheap(r) = \codereg{\sigrets,\sigcloss,\mscode}$, and either $[\baddr,\eaddr] \subseteq \ta$ and $\mscode$ behaves reasonably (see Section~\ref{sec:well-form-reas}) or
 $[\baddr,\eaddr] \mathrel{\#} \ta$,
@@
 \end{lemma}
-Roughly speaking, the FTLR says: any executable capability used for execution in a safe environment won't break the capability machine guarantees.
-This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ entails, and for trusted code the semantic reasonability condition must also be satisfied.
-The proof of the FTLR is done by a nested induction where each instruction that could be executed is proven safe.
-The lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, but it is crucial to the proofs of the lemmas in that section.
+
+Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
+The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see \todo{explain}).
+
+The proof of the lemma essentially consists of a big induction where each possible instruction is proven to behave the same in source and target in related memories and register files.
+After that first step, the induction hypothesis is used for the rest of the execution.
 
@@
 \label{subsec:proof-sketch}
-The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straightforward manner.
-The component relation relates components to themselves which makes sense as well-formed components should contain no \srccm{} specific constructs.
-We prove a number of lemmas about the relations.
-The first is a FTLR for components that says (1) all well-formed untrusted components are related to themselves in the component relation, and (2)  all well-formed and reasonable trusted components are related to themselves in the component relation.
-% \subsection{Component logical relation}
+Using Lemma~\ref{thm:ftlr}, we can now proceed to proving Theorem~\ref{thm:full-abstraction} (full abstraction).
+First, we will extend the logical relation into an omitted relation on components $(\var{comp}_S,\var{comp}_T) \in \lrcomp(W)$.
+Essentially, it relates a component to itself if instantiating its imports with related values produces related exports and code memory satisfying the appropriate code region in the world $W$.
+
 % The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
@@
 % \end{multline*}
+
+Using Lemma~\ref{thm:ftlr} and the definitions of the logical relations, we can then prove the following two lemmas.
+The first is a version of the FTLR for components, stating that all components are related to themselves if they are either (1) well-formed and untrusted or (2) well-formed, reasonable and trusted.
 \begin{lemma}[FTLR for components]
@@
 \end{lemma}
-The next is a compatibility lemma for plugging that says: given a context and component each related to itself in the component relation, plugging the context with the component will give an executable configuration in the executable configuration relation.
-\begin{lemma}[Compatibility lemma for context plugging]
-  \label{lem:compat-context-plug}
-  If $\npair{\stpair{\context}{\context}}\in\lrcomp(W_1)$ and $\npair{\stpair{\comp}{\comp}} \in \lrcomp(W_2)$, then
-  $\npair[n']{(\plug{\src{\context_S}}{\src{\comp_S}},\plug{\context_T}{\comp_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.
+
+Another lemma then relates the component relation and context plugging: given a context and component each related to themselves in the component relation, plugging the context with the component produces an executable configuration in the executable configuration relation.
+\begin{lemma}
+  \label{lem:adeq-context-plug}
+  If $\npair{\stpair{\context}{\context}}\in\lrcomp(W_1)$ and $\npair{\stpair{\comp}{\comp}} \in \lrcomp(W_2)$ and $W_1\oplus W_2$ is defined, then
+  $\plug{\src{\context_S}}{\src{\comp_S}}$ terminates iff $\plug{\context_T}{\comp_T}$ terminates.
 \end{lemma}
-Finally, we have an adequacy lemma for the execution configuration relation.
-This lemma says that if an \srccm{} configuration approximates a \trgcm{} configuration, and the \trgcm{} configuration terminates, then so does the \srccm{} configuration.
-The symmetric case is also true.
+% Finally, we have an adequacy lemma for the execution configuration relation.
+% This lemma says that for related \srccm{} and \trgcm{} configurations, one terminates iff the other does.
 % \begin{multline*}
@@
 % \end{multline*}
-\begin{lemma}[Adequacy of execution configuration logical relation]
-  \label{lem:adequacy}
-  If $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\preceq,\gc](W)$ and
- $i \leq n$ and
-and $\src{\Phi_S} \sterm[i]{\gc}$,
-  then $\Phi_T\term$.
-
-  Also, if
- $\npair{\stpair{\Phi}{\Phi}}\in\lrec[\succeq,\gc](W)$ and
- $i \leq n$
-and $\Phi_T \term[i]$,
-  then $\src{\Phi_S}\sterm{\gc}$.
-\end{lemma}
-We use these three lemmas to prove Theorem~\ref{thm:full-abstraction}
+% \begin{lemma}[Adequacy of execution configuration logical relation]
+%   \label{lem:adequacy}
+%   If $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W)$ then $\src{\Phi_S} \sterm{}$ iff $\Phi_T\term$.
+% \end{lemma}
+
+Finally, we use these two lemmas to prove Theorem~\ref{thm:full-abstraction}
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
   The proof of both directions are very similar, so we only show the right direction.
-  We need to show the \trgcm{} contextual equivalence, so w.l.o.g\ assume ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
+  We need to show the \trgcm{} contextual equivalence, so w.l.o.g\ assume a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
   The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.
   By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that the trusted components $\src{\comp_1}$ and $\src{\comp_2}$ are well-formed and reasonable.
-  Further, by $\tconeq$ we may assume the context $\trg{\context}$ is well-formed.
-  We prove arrow (1) in the figure by using the mentioned assumptions about $\src{\comp_1}$ and $\trg{\context}$ along with Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}.
-  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$, so by the assumption that $\comp_1$ and $\comp_2$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$, i.e.\ arrow (2) in the figure.
-  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:compat-context-plug}, and \ref{lem:adequacy}; but this time, we use the assumption that $\src{\comp_2}$ is well-formed and reasonable and that $\trg{\context}$ is well-formed.
+  We prove arrow (1) in the figure by using the mentioned assumptions about $\src{\comp_1}$ and $\trg{\context}$ along with Lemma~\ref{lem:ftlr-comps} and \ref{lem:adeq-context-plug}.
+  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{}{}}$, so by the assumption that $\src{\comp_1}$ and $\src{\comp_2}$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{}{}}$, i.e.\ arrow (2) in the figure.
+  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:adeq-context-plug}; but this time, we use the assumption that $\src{\comp_2}$ is well-formed and reasonable and that $\trg{\context}$ is well-formed.
 \end{proof}

### 2018-10-16 fully-abstract-comp/popl2019/paper.tex
Rewrite the start of section 4 to explain our formulation of WBCF and LSE.

@@
 
-\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{\gc}}
+\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
 \renewcommand{\stdreg}[3][]{\iota^{\mathrm{std},#3,#1}_{#2}}
@@
 
-A second contribution of this work is the way in which we formulate these two properties using a technique we call {\itshape fully abstract overlay semantics}.
+A second contribution of this work is the way in which we formulate these two properties: although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
 Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
+We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
 Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
@@
 \label{sec:form-secur-with}
-A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
-Our proof uses a novel approach we call fully abstract overlay semantics.
+As mentioned, the \stktokens{} calling convention guarantees well-bracketed control flow (WBCF) and local state encapsulation (LSE).
+However, before we can prove these properties, we need to know how to even formulate them.
+Although the properties are intuitively clear and sound precise, formalizing them is actually far from obvious.
+
+Ideally, we would like to define the properties in a way that is
+\begin{enumerate}
+\item {\itshape intuitive}
+\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}.
+\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when reasoning about correctness and security of other compiler passes and then compose such results with our results to obtain results about the full compiler.
+\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
+\item {\itshape potentially scalable}: even though we currently do not consider dynamic code
+  generation or multi-threading, the current formalization should at least potentially extend to such settings.
+\end{enumerate}
+
+Previous formalisations in the literature are formulated in terms of a static control flow graph~\cite[e.g., ]{Abadi2005Theory}.
+While these are intuitively appealing (1), it is not clear how they can be used to reason about programs (2) or other compiler passes (3), they lack temporal safety guarantees (4) and do not scale (5) to settings with dynamic code generation (where a static control flow graph cannot be defined).
+Skorstengaard et al.
+[2018] provide a logical relation that can be used to reason about programs using their calling convention (2,3), but it is not intuitive (1), there is no argument for completeness and it is unclear whether it will scale to more complex features.
+
+We contribute a new way to formalise the properties using a novel approach we call fully abstract overlay semantics.
 The idea is to define a second operational semantics for programs in our target language.
@@
 Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
-The most important relations are summarised in the following table:\\
+The most important relations are summarised in the following table, where we mention the general form of the relations, what type of things they relate and extra conditions that some of them imply:\\
 % latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
@@
   \hline
-  \npair{(w_S,w_T)} & \lrv(W) & values & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\
-  \npair{(w_S,w_T)} & \lrvtrusted(W) & values & \\
+  \npair{(w_S,w_T)} & \lrv(W) & values (machine words) & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\
+  \npair{(w_S,w_T)} & \lrvtrusted(W) & values (machine words) & \\
   \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\

### 2018-10-16 fully-abstract-comp/popl2019/paper.tex
Comments and corrections in Section 5

@@
 They consist of three sub-worlds: $\Wor = \Worldh \times \Worlds \times \Worldfs$, capturing assumptions about the heap\footnote{Actually, we use the term heap to describe all memory except the stack, including, for example, code memory.}, the inactive and the active part of the stack, respectively.
-Sub-worlds consist of a finite mapping from region names to regions, which come in two forms (spatial and shared):
+Sub-worlds consist of a finite mapping from region names to regions which come in two forms (spatial and shared):
 \begin{align*}
@@
 \end{align*}
-Different parts of the world can contain different types of regions: heap-related assumptions can be either shared or spatial (see below), while stack-related regions must be spatial.
+Different parts of the world can contain different types of regions: heap-related assumptions can be either spatial or shared (see below), while stack-related regions must be spatial.
 Additionally, regions for inactive parts of the stack additionally include an address specifying the return address for that stack frame.
@@
 Shared regions are tagged with the syntactic token $\pure$ and may prescribe two different types of requirements.
-First, they may require the presence of \srccm{} and \trgcm{} memory segments satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$ (see below for the meaning of $\Wor \monnefun$).
+First, they may require the presence of \srccm{} and \trgcm{} memory segments
+satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$
+% Lau: Where do you return to this? I did not find it very clear what it meant when I was done reading about the worlds.
+(see below for the meaning of $\Wor \monnefun$).
 A region might, for example, require the presence of a certain list of instructions at a given set of memory addresses in both \srccm{} and \trgcm{}.
 If a memory segment is owned by a given region, then that memory segment must be disjoint from the memory segments owned by all other regions.
-Second, a shared region may also contains a partial function from seals to relations on pairs of sealable capabilities: $\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps}$ (see below for the meaning of $\Wor \monnefun$).
+Second, a shared region may also contain a partial function from seals to relations on pairs of sealable capabilities: $\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps}$.
 When the region defines such a relation for a given seal, then no other region in the world can do the same, and any value signed with that seal will be required to satisfy the registered relation.
@@
 For example, a \trgcm{} linear capability to a piece of memory is related to its \srccm{} counterpart, but only if no other linear capability overlaps with it.
-We will model such an assumption for exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region, that only one value in the system may rely on.
+We will model such an assumption for exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region that only one value in the system may rely on.
 More concretely, we will define how to combine worlds $W_1$ and $W_2$ into a combined world $W_1 \oplus W_2$, on the condition that they represent compatible assumptions: $W_1$ and $W_2$ must contain the same regions except that they must respect exclusive ownership of $\spatialo$ regions: $\pure$ regions must be present in both worlds, but a $\spatialo$ region can only be present in one and must be $\spatial$ in the other.
@@
 
-Attentive readers may have noticed that our definition of worlds is actually cyclic: worlds in $\Wor$ contain regions in $Regionh$, but those contain partial functions from the set $\Wor$ to something else.
+Attentive readers may have noticed that our definition of worlds is actually cyclic: worlds in $\Wor$ contain regions in $\Regionh$, but those contain partial functions from the set $\Wor$ to something else.
 Such recursive worlds are in fact common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} (essentially an advanced form of step-indexing) to construct the set $\Wor$ and rigorously resolve the circularity.
@@
 First, we have the code region $\codereg{\sigrets,\sigcloss,\mscode}$.
-This is a shared region that represents the assumption that memory segment $\mscode$ is loaded in heap memory at a certain location, that it is well-formed and uses return seals $\sigrets$ and closure seals $\sigcloss$.
+This is a shared region that represents the assumption that memory segment
+$\mscode$ is loaded in heap memory at a certain location, it is well-formed and it uses return seals $\sigrets$ and closure seals $\sigcloss$.
+% Lau: I cannot parse this sentence:
 It takes ownership of those seals and registers appropriate invariants on the capabilities that may be signed with them.
@@
 A second type of regions $\stdreg{A}{\pure}$ and $\stdreg{A}{\spatialo}$ governs heap or stack memory at a set of addresses $A$.
-It simply requires presence of a memory segment for those addresses, such that the memory contains values that themselves satisfy the relation between values that we will see below.
+It simply requires the presence of a memory segment for those addresses, such that the memory contains values that themselves satisfy the relation between values that we will see below.
 The contents of the memory is allowed to change as long as the new contents are still valid.
-Finally, a third type of regions $\stareg[\stpair{\ms}{\ms}]{\pure}$ or $\stareg[\stpair{\ms}{\ms}]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
+Finally, a third type of regions $\stareg[(\ms_S,\ms_T)]{\pure}$ or $\stareg[(\ms_S,\ms_T)]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
 This region is used, for example, to govern inactive parts of the stack whose contents is required to remain unmodified.
@@
 % latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
+% Lau: I find it difficult to easily see what constitutes a row in this table.
+% I have updated it to a table I find easier to read.
 \newcolumntype{R}{>{$}r<{$}}
@@
 \newcolumntype{M}{R@{${}\in{}$}L}
-\begin{tabular}{|M|c|c|}
+\begin{tabular}{|M|c|p{4.8cm}|}
   \hline
@@
   \hline
-  \npair{(w_S,w_T)} & \lrv(W) & values (machine words) & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\
+  \npair{(w_S,w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
   \npair{(w_S,w_T)} & \lrvtrusted(W) & values (machine words) & \\
-  \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & \parbox{3.3cm}{safe to pass to\\ adversarial code}\\
+  \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
   \npair{(\reg_S,\reg_T)}  &  \lrrtrusted(W) & register files & \\
@@
   \npair{(w_S,w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
-  \npair{((w_{S,1},w_{S,2}),(w_{T,1},w_{T,2}))}  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
-  \multicolumn{2}{|c|}{$\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$} & memory & \parbox{3.3cm}{satisfy the\\ assumptions in $W$}\\
+  \left(\arraycolsep=1pt\array{l}(w_{S,1},w_{S,2}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
+  \multicolumn{2}{|c|}{$\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
   \hline
@@
 These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another (omitted) use of step-indexing.
-For space reasons, we cannot show all of these definitions, but we will try to give a good overview, omitting many details.
+For space reasons, we cannot show all of these definitions, but we will try to give a good overview.
 
@@
     \begin{array}{l}
+    % Lau: Consider using \Wor instead of \World as we have not made a distinction between the two.
       \exists S : (\RegName \setminus \{\pcreg \})\fun \World \ldotp \\

### 2018-10-16 fully-abstract-comp/popl2019/paper.tex
Minor simplifications and fixes in the new callCondition and friends.
Proposed re-refactoring of the call semantics.

--- 2018-10-16 fully-abstract-comp/popl2019/paper.tex
@@
   \end{mathpar}
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.8cm}<{$}|>{\raggedright\arraybackslash}p{6.4cm}|}
-    \hline
-    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
-    \hline
-    \halt                                        & \halted & \\
+  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.55cm}|}
+    \hline
+    i \in \Instr                                 & $\sem{i}(\Phi)$ & Conditions\\
+    \hline
+    \halt                                        & $\halted$ & \\
     \hline
@@
     \hline
-    \store{r_1}{r_2}                             & \srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
-    \hline
-    \cca{r}{\rn}                                 &\srcalt{\updPcAddr{\Phi\updReg{r}{w}}} &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
-    \hline
-    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} & \srcalt{\var{xjmpRes}(c_1,c_2,\Phi')} & $\srcalt{ \Phi' =\var{callConditions}(r_1,r_2,\offpc,\offsigma,\Phi)}$\\
+    \store{r_1}{r_2}                             & $\srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray$  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
+    \hline
+    \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
+    \hline
+    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} &
+        $\srcalt{\var{xjmpRes}(c_1,c_2,}$
+        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1,r_2}{w_1,w_2}\\
+            &\updReg{\rretc}{s_c}\\
+            &\updReg{\rretd}{s_d}\\
+            &\updReg{\rstk}{c_\stk}\\
+            &\update{\ms_\stk}{\ms_{\stk,\var{rest}}}\\
+            &\update{\stk}{\stk'}\\
+            \endarray\right))}$
+       &
+      \srcalt{$\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
+         $\begin{multlined}
+           \ms_{\stk,\var{local}},\ms_{\stk,\var{rest}},c_\stk,c_\var{local} =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
+         \end{multlined}$
+    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
+    $\sigma = \var{getCallSeal}(\Phi.\reg,\Phi.\mem,\offpc,\offsigma)$ and 
+    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local},c_\opc)$
+    $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
+                        $\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
+                        }\\
     \hline 
@@
     \hline
-    \_                                           & \failed & \totherwise \\
+    \_                                           & $\failed$ & \totherwise \\
     \hline
@@
 \end{multline*}
-  \caption{An excerpt of the operational semantics of \srccm{}.}
+  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted).}
   \label{fig:source-op-sem}
@@
     \[
-      \src{ \opc, c_\opc = \var{setupOpc(\reg)}} \mathit{ iff }
+      \src{ \opc, c_\opc = \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} \mathit{ iff }
       \left\{ 
       \begin{array}{l}
-      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge{}}\\
       \src{\opc = \aaddr + \calllen \wedge{}}\\
@@
     \begin{multline*}
-      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk)} \mathit{\ iff} \\
+      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff} \\
       \left\{ 
       \begin{array}{l}
-      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \wedge \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
+      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
       \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
-      \src{\ms_{\stk,\var{unused}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} }\\
-      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }
+      \src{\ms_{\stk,\var{unused}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]} }\\
+      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
       \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
@@
     \end{multline*}
-    \[
-      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}})}\text{\ iff\ }
-            \src{\stk' = (\opc,\ms_{\stk,\priv}) :: \stk }
-    \]
     \[
@@
     \]
-  \begin{multline*}
-    \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}\mathit{\ iff}\\
-      \quad\quad\left\{ 
-        \begin{array}{l}
-          \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)}\\
-      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{setupStackAndDataRetPtr}(\reg, \ms_\stk) \wedge{}}\\
-      \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
-      \src{\stk' = \var{pushCallStackFrame}(\stk, \opc, \ms_{\stk,\var{unused}}) \wedge{}}\\
-      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
-      \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc) \wedge{}}\\
-      \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
-      \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
-      \end{array}\right.
-  \end{multline*}
+  % \begin{multline*}
+  %   \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}\mathit{\ iff}\\
+  %     \quad\quad\left\{ 
+  %       \begin{array}{l}
+  %         \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)}\\
+  %     \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{rest}},c_\stk,c_\var{local\_data} = \var{splitStack}(\reg, \ms_\stk) \wedge{}}\\
+  %     \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
+  %     \src{\stk' = (\opc,\ms_{\stk,\var{local}}) :: \stk \wedge{}}\\
+  %     \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
+  %     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local\_data},c_\opc) \wedge{}}\\
+  %     \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
+  %     \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
+  %     \end{array}\right.
+  % \end{multline*}
 

### 2018-10-17 fully-abstract-comp/popl2019/paper.tex
phase-2 version should no longer be anonymous
More work on figure 10.

@@
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
          $\begin{multlined}
-           \ms_{\stk,\var{local}},\ms_{\stk,\var{rest}},c_\stk,c_\var{local} =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
+           \ms_{\stk,\var{local}},c_\var{local},\ms_{\stk,\var{rest}},c_\stk =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
          \end{multlined}$
@@
     $\sigma = \var{getCallSeal}(\Phi.\reg,\Phi.\mem,\offpc,\offsigma)$ and 
-    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local},c_\opc)$
+    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local},c_\opc)$ and
     $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
-                        $\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
+                        $\Phi.\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
                         }\\
@@
     &\phantom{mak} & $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
-        \src{\Phi'} 
+      \arraycolsep=0pt
+      \array[c]{rl}
+      \Phi&\updReg{\pcreg}{c_\opc}\\
+          &\updReg{\rstk}{c_\stk}\\
+          &\updReg{\rdata}{0}\\
+          &\update{\stk}{\stk'}\\
+          &\update{\ms_\stk}{\ms_\stk \uplus \ms_\var{local}}
+      \endarray
            & &
-        $\src{ \Phi' = \var{returnCondition}(c_1,c_2,\Phi)}$ \\
+               $\arraycolsep=0pt\array{l}
+        \src{(\opc, \ms_\var{local}) :: \stk'= \Phi.\stk \tand} \\
+      \src{\var{retPtrMatchesOpc}(c_1,\opc)}\\
+      \src{\var{retPtrForLocalStack}(c_2,\ms_\var{local})}\\
+      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2) \tand}\\
+      \src{c_\opc =  \var{tokenToPc}(c_1)} 
+      \endarray$\\
       \failed &  & \totherwise
@@
     \begin{multline*}
-      \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{unused}},c_\stk,c_\var{local\_data} = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff} \\
+      \src{\ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff} \\
       \left\{ 
@@
 \]
+% \[
+%     \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
+%     \mathit{\ iff\ }
+%       \left\{ \array{l}
+%         \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
+%         \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
+%         \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
+%         \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
+%         \endarray
+%      \right.
+% \]
 \[
-    \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
-    \mathit{\ iff\ }
-      \left\{ \array{l}
-        \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
-        \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
-        \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
-        \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
-        \endarray
-     \right.
+  \src{\var{retPtrMatchesOpc(\retptrc(\_,\_,\opc_c),\opc)}} \mathit{\ iff\ } \src{\opc_c = \opc}
 \]
+\[
+  \src{\var{retPtrForMatchLocalStack(\retptrd(\aaddr_\stk,\eaddr_\stk),\ms_\var{local})}} \mathit{\ iff\ } \src{\dom(\ms_\var{local} \subseteq [\aaddr_\stk,\eaddr_\stk])}
+\]
+
 \[
@@
 
-\begin{multline*}
-    \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
-      \quad\quad\left\{ 
-      \begin{array}{l}
-      \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
-      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
-      \src{c_\opc =  \var{tokenToPc}(c_1)} \\
-      \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
-      \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
-      \end{array}\right.
-  \end{multline*}
+% \begin{multline*}
+%     \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
+%       \quad\quad\left\{ 
+%       \begin{array}{l}
+%       \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
+%       \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
+%       \src{c_\opc =  \var{tokenToPc}(c_1)} \\
+%       \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
+%       \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
+%       \end{array}\right.
+%   \end{multline*}
 

### 2018-10-18 fully-abstract-comp/popl2019/paper.tex
Clarify attacker model a bit

@@
 However, \stktokens{} is a modular calling scheme: no restriction is put on adversarial components that we do not expect trusted components to meet.
+Specifically, we are going to assume that both trusted and adversarial components are initially syntactically well-formed (described in more detail in Section~\ref{sec:well-form-reas}) which basically just restrict adversarial components to not break machine guarantees initially (e.g.\ no aliases for linear capabilities or access to seals of other components).
 This means that any component can ensure WBCF and LSE by employing \stktokens{}.
-% KJAA: ^-- Assuming that there is a global stack address capability, no? Can I "suddenly" begin to impose \stktokens{}?
+
+% KJAA: ^-- Assuming that there is a global stack address capability, no?
+Can I "suddenly" begin to impose \stktokens{}?
 
@@
 Our formal results are complicated by the fact that they only hold for well-behaved components that respect the basic rules of the calling convention, and also on a sane initial configuration of the system.
-For example, the system should be set up such that seals that components use for constructing return pointers are not shared with other components.
+For example, the system should be set up such that seals used by components for constructing return pointers are not shared with other components.
 We envision distributing seals as a job for the linker, so this means our results depend on a linker to do this properly.
-As another example, a seal that is used to construct a return pointer can be used again to construct other return pointers for the same return point, but a different seal must be used for others.
+As another example, a seal used to construct a return pointer can be reused but only to construct return pointers for the same return point.
+Different seals must be used for different return points.
 Such seals should also never be passed to other components.
@@
 % - Address must be in data memory
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the syntactic requirements necessary to be able to rely on system guarantees.
-We elide the details of the judgement here and describe the main points.
+The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
+We elide the details of the judgement here and describe the main points:
 For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
@@
 The return seals $\sigrets$ are the seals supposed to be used to seal return pointers and the closure seals $\sigcloss$ all the other seals in a component .
-If a component is not trusted, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
-The code memory may contain sets of seals, but only the seals in $\sigrets$ and $\sigcloss$.
+If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
+The code memory may contain sets of seals but only the seals in $\sigrets$ and $\sigcloss$.
 Other than that, code memory only contains instructions in the form of integers.
-When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), it must have access to the return seal specified by that call.
-The return seal also needs to be unique to that call, so no other call can specify the same seal.
+When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), the call must have access to the return seal it specifies.
+The return seal also needs to be unique to that call; that is, no other call can specify the same seal as its return seal.
 The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
 The capabilities in the data memory can only have authority over the data memory itself.
-By allowing this, a component can be constructed with predetermined data structures.
-In order to respect Write-XOR-Execute, the capabilities can at most have a read-write permission.
-The capabilities also need to respect linearity which means that a linear capability cannot have any aliases, and it must be for a range of the predetermined linear addresses $A_\linear$.
+This allows components to have initial data structures.
+In order to respect Write-XOR-Execute, the capabilities in data memory cannot have execute permission.
+The capabilities also need to respect linearity which means that linear capabilities cannot be aliased by any other capability, and they must be for a range of the predetermined linear addresses $A_\linear$.
 The sealed capabilities must be sealed with a closure seal, and the sealable can be anything that is allowed to reside in data memory.
 The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
-Finally, the addresses of the imports $\overline{\var{import}}$ must be addresses in the data memory.
 
@@
 We elide the details here and refer to~\citet{technical_report}
+
+In our result, we assume that adversarial components are well-formed, but not necessarily reasonable.
+The well-formedness assumption ensures that the trusted component can rely on basic security guarantees provided by the capability machine.
+For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
+The adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
+Can adversarial code call into trusted components?
+The answer to that question is yes but not with LSE and WBCF guaranteed.
+Formally, adversarial code can contain the instructions that constitute a call.
+However, for untrusted code, \srccm{} will not execute those instructions as a "native call" but execute the individual instructions separately.
+The callee then executes in the same stack frame as the caller, so WBCF and LSE do not follow (for that call).
+
+We will assume trusted component, for which WBCF and LSE are guaranteed, to be both well-formed and reasonable. 
 %The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.

### 2018-10-18 fully-abstract-comp/popl2019/paper.tex
Move permission hierarchy figure
Various small improvements

@@
 %%% Capabilities
-\begin{wrapfigure}{r}{0.26\linewidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (rwx) {$\rwx$};
-    \node[main node] (rx) [below right of=rwx] {$\rx$};
-
-    \node[main node] (rw) [below left of=rwx,] {$\rw$};
-    \node[main node] (r) [below right of=rw] {$\readonly$};
-    \node[main node] (0) [below of=r] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-
-    (rw) edge (r)
-    (r) edge (0)
-
-    (rwx) edge (rx)
-
-    (rw) edge (rwx)
-    (r) edge (rx);
-  \end{tikzpicture}
-
-  \caption{Permission hierarchy}
-  \label{fig:perm-hier}
-\end{wrapfigure}
 The concept of a capability is the cornerstone of any capability machine.
@@
 The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
+\begin{wrapfigure}{r}{0.26\linewidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (rwx) {$\rwx$};
+    \node[main node] (rx) [below right of=rwx] {$\rx$};
+
+    \node[main node] (rw) [below left of=rwx,] {$\rw$};
+    \node[main node] (r) [below right of=rw] {$\readonly$};
+    \node[main node] (0) [below of=r] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+
+    (rw) edge (r)
+    (r) edge (0)
+
+    (rwx) edge (rx)
+
+    (rw) edge (rwx)
+    (r) edge (rx);
+  \end{tikzpicture}
+
+  \caption{Permission hierarchy}
+  \label{fig:perm-hier}
+\end{wrapfigure}
 
@@
               \and
-              \inference{ \forall \Phi' \ldotp \neg (\Phi \step \Phi') \wedge \Phi' \neq \failed}
+              \inference{ \forall \Phi' \neq \failed \ldotp \Phi \nrightarrow \Phi'}
                         {\Phi \step \failed }
@@
 When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
-In the case of \texttt{move}, the source register $r_2$ is overwritten with $\linCons{\Phi(r_2)}$.
+In the case of \texttt{move}, the source register $\rn$ is overwritten with $\linCons{\Phi(\rn)}$.
 We see the other approach later.
@@
 The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
-\texttt{cca} changes the current address or seal of a capability or set of seals, respectively, by a given offset.
+\texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
 It handles linearity by updating the capability in-place, i.e.\ the source register is used as the target register.
@@
 The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
-To make this clearer, we introduce notions of components and programs from which we construct executable configurations.
+To clarify this, we introduce notions of components and programs from which we construct executable configurations.
 A component is basically a program with entry points in the form of imports that need to be linked.
@@
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
+As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not seem to provide sufficient guarantees to enforce LSE and WBCF.
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
@@
 A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
-However without any enforcement, there is nothing to prevent a callee from returning on an arbitrary call on the call stack.
+However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
 This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
@@
 On return the callee has to give back the stack capability they were given on invocation.
-When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack.
+When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack or not.
 They do this by attempting to restore the stack capability from before the call by splicing the return token with the stack capability for the local stack frame which at this point has been unsealed again.
@@
 On the other hand, if the splice fails, then they are alerted to the fact that their stack frame may not be the topmost.
-\stktokens{} uses this approach, and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev} it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
+\stktokens{} uses this approach; and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}, it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
 \begin{figure}
@@
 \begin{enumerate}
-\item \label{item:check-stkb} Check the base address of the stack capability before and after calls
-\item \label{item:non-empty-sf} Make sure that local stack frames are non-empty
+\item \label{item:check-stkb} Check the base address of the stack capability before and after calls.
+\item \label{item:non-empty-sf} Make sure that local stack frames are non-empty.
 \item \label{item:return-data} Create token and data return capability on call: split the stack capability in two to get a stack capability for your local stack frame and a stack capability for the unused part of the stack.

### 2018-10-20 fully-abstract-comp/popl2019/paper.tex
Propose some re-inlinings in Fig.10, 11, 12.
Further minor refactoring in getCallSeal.
Reduce the crammedness of Fig. 10 with some latex relayouting...
Further Figure 10 refactoring
- Changed the order of sealReturnPair arguments to be the same order as result.
- Changed c_priv_stk to c_local in aux definition for consistency.
- Formatting
Removed sealedPairArgument aux definition
It looks like this definition had become obsolete.
Reorder aux definitions to match the order in definitions
Put scall on one line

@@
   \end{mathpar}
-  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.55cm}|}
+  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.6cm}|}
     \hline
@@
     \hline
-    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}} &
+    \arraycolsep=0pt\array{c}
+    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}}
+    \endarray &
         $\srcalt{\var{xjmpRes}(c_1,c_2,}$
@@
     $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
-    $\sigma = \var{getCallSeal}(\Phi.\reg,\Phi.\mem,\offpc,\offsigma)$ and 
-    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local},c_\opc)$ and
+    $\sigma =\begin{multlined}[t]
+      \var{getCallSeal}(\\
+      \Phi.\reg(\pcreg), \Phi.\mem,\offpc,\offsigma)  \text{ and }
+    \end{multlined}$
+    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\opc,c_\var{local})$ and
     $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
@@
         \src{(\opc, \ms_\var{local}) :: \stk'= \Phi.\stk \tand} \\
-      \src{\var{retPtrMatchesOpc}(c_1,\opc)}\\
-      \src{\var{retPtrForLocalStack}(c_2,\ms_\var{local})}\\
-      \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2) \tand}\\
-      \src{c_\opc =  \var{tokenToPc}(c_1)} 
+      \src{c_1 = \retptrc(\baddr,\eaddr,\opc)}\\
+      \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge \dom(\ms_\var{local}) \subseteq [\aaddr_\stk,\eaddr_\stk]}\\
+      \src{c_\stk = \var{reconstructStackPointer}(\reg(\rstk),c_2) \tand}\\
+      \src{c_\opc =  ((\rx,\normal),\baddr,\eaddr,\opc)} 
       \endarray$\\
@@
 \begin{figure}
-    \[
-      \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)} \mathit{ iff }
-      \left\{ 
-      \begin{array}{l}
-        \src{\exists \sigma \ldotp (\reg(r_1) = \sealed{\sigma, c_1} \wedge \reg(r_2) = \sealed{\sigma, c_2}) \wedge{}}\\
-        \src{r_1 \neq \rtmp{1} \wedge r_2 \neq \rtmp{2}}
-      \end{array}
-      \right.
-    \]
-    \[
-      \src{ \opc, c_\opc = \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} \mathit{ iff }
-      \left\{ 
-      \begin{array}{l}
-      \src{\opc = \aaddr + \calllen \wedge{}}\\
-      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
-      \end{array}
-       \right.
-    \]
     \begin{multline*}
-      \src{\ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff} \\
+      \src{\ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff\ } \\
       \left\{ 
@@
     \[
-      \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma)} \mathit{\ iff\ }
+      \src{ \opc, c_\opc = \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} \mathit{\ iff\ }
       \left\{ 
       \begin{array}{l}
-      \src{\reg(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
+      \src{\opc = \aaddr + \calllen \wedge{}}\\
+      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
+      \end{array}
+       \right.
+    \]
+    \[
+      \src{\sigma = \var{getCallSeal}(c_\pcreg,\mem,\offpc,\offsigma)} \mathit{\ iff\ }
+      \left\{
+      \begin{array}{l} 
+      \src{c_\pcreg = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
       \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge{} \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \wedge{}}\\
@@
     \[
-     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{priv\_data},c_\opc)} \mathit{\ iff\ }
-      \src{s_d = \sealed{\sigma,c_\var{priv\_data}} \wedge{} s_c = \sealed{\sigma,c_\opc}}
+     \src{\sealed{\sigma,c_\opc},\sealed{\sigma,c_\var{local}} = \var{sealReturnPair}(\sigma,c_\opc,c_\var{local})}
     \]
@@
 
-  \caption{The call condition in the operational semantics of \srccm{}.}
+  \begin{multline*}
+    \src{\var{reconstructStackPointer}(\stkptr{\rw, \stkb, \aaddr_\stk-1,\_},\retptrd(\aaddr_\stk,\eaddr_\stk)) =}\\ \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk}} \mathit{\ iff\ }
+        \src{\stkb \leq \addr_\stk }
+   \end{multline*}
+
+  \caption{Auxiliary definitions used in the operational semantics of \srccm{}.}
   \label{fig:source-op-sem}
 \end{figure}
-\begin{figure}
-\[
-    \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
-\]
-\[
-    \src{((\rx,\normal),\baddr,\eaddr,\opc) = \var{tokenToPc}(\retptrc(\baddr,\eaddr,\opc)}
-\]
-% \[
-%     \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
-%     \mathit{\ iff\ }
-%       \left\{ \array{l}
-%         \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
-%         \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
-%         \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
-%         \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
-%         \endarray
-%      \right.
-% \]
-\[
-  \src{\var{retPtrMatchesOpc(\retptrc(\_,\_,\opc_c),\opc)}} \mathit{\ iff\ } \src{\opc_c = \opc}
-\]
-\[
-  \src{\var{retPtrForMatchLocalStack(\retptrd(\aaddr_\stk,\eaddr_\stk),\ms_\var{local})}} \mathit{\ iff\ } \src{\dom(\ms_\var{local} \subseteq [\aaddr_\stk,\eaddr_\stk])}
-\]
-
-\[
-    \src{c_\stk = \var{reconstructStackPointer}(\reg,c)}
-    \mathit{\ iff\ }
-      \left\{ \array{l}
-        \src{c = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge{}} \\
-        \src{\reg(\rstk) = \stkptr{\rw, \stkb, \aaddr_\stk-1,\_} \wedge{}} \\
-        \src{\stkb \leq \addr_\stk \wedge{}} \\
-        \src{c_\stk = \var{tokenToStkPtr(c)} }
-        \endarray
-     \right.
-\]
-
-% \begin{multline*}
-%     \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
-%       \quad\quad\left\{ 
-%       \begin{array}{l}
-%       \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
-%       \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
-%       \src{c_\opc =  \var{tokenToPc}(c_1)} \\
-%       \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
-%       \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
-%       \end{array}\right.
-%   \end{multline*}
-
-  \caption{The return condition in the operational semantics of \srccm{}.}
-  \label{fig:source-op-sem-return}
-\end{figure}
+% \begin{figure}
+% % \[
+% %     \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
+% % \]
+% % \[
+% %     \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
+% %     \mathit{\ iff\ }
+% %       \left\{ \array{l}
+% %         \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
+% %         \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
+% %         \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
+% %         \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
+% %         \endarray
+% %      \right.
+% % \]
+% % \[
+% %   \src{\var{retPtrForMatchLocalStack(\retptrd(\aaddr_\stk,\eaddr_\stk),\ms_\var{local})}} \mathit{\ iff\ } \src{\dom(\ms_\var{local} \subseteq [\aaddr_\stk,\eaddr_\stk])}
+% % \]
+
+% % \begin{multline*}
+% %     \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
+% %       \quad\quad\left\{ 
+% %       \begin{array}{l}
+% %       \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
+% %       \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
+% %       \src{c_\opc =  \var{tokenToPc}(c_1)} \\
+% %       \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
+% %       \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
+% %       \end{array}\right.
+% %   \end{multline*}
+
+%   \caption{The return condition in the operational semantics of \srccm{}.}
+%   \label{fig:source-op-sem-return}
+% \end{figure}
 \subsection{Well-formedness and reasonability}

### 2018-10-20 fully-abstract-comp/popl2019/paper.tex
Update WBCF and LSE propety part to use LaTeX references

@@
 \begin{enumerate}
-\item {\itshape intuitive}
-\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}.
+\item {\itshape intuitive} \label{def-prop:intuitive}
+\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
 \item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when reasoning about correctness and security of other compiler passes and then compose such results with our results to obtain results about the full compiler.
+\label{def-prop:reusable}
 \item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
-\item {\itshape potentially scalable}: even though we currently do not consider dynamic code
-  generation or multi-threading, the current formalization should at least potentially extend to such settings.
+\label{def-prop:complete}
+\item {\itshape potentially scalable}: even though we currently do not consider dynamic code generation or multi-threading, the current formalization should at least potentially extend to such settings.
+\label{def-prop:scalable}
 \end{enumerate}
@@
 Previous formalisations in the literature are formulated in terms of a static control flow graph~\cite[e.g., ]{Abadi2005Theory}.
-While these are intuitively appealing (1), it is not clear how they can be used to reason about programs (2) or other compiler passes (3), they lack temporal safety guarantees (4) and do not scale (5) to settings with dynamic code generation (where a static control flow graph cannot be defined).
-Skorstengaard et al.
-[2018] provide a logical relation that can be used to reason about programs using their calling convention (2,3), but it is not intuitive (1), there is no argument for completeness and it is unclear whether it will scale to more complex features.
+While these are intuitively appealing (\ref{def-prop:intuitive}), it is not clear how they can be used to reason about programs (\ref{def-prop:useful}) or other compiler passes (\ref{def-prop:reusable}), they lack temporal safety guarantees (\ref{def-prop:complete}) and do not scale (\ref{def-prop:scalable}) to settings with dynamic code generation (where a static control flow graph cannot be defined).
+\citet{skorstengaard_reasoning_2017} provide a logical relation that can be used to reason about programs using their calling convention (\ref{def-prop:useful},\ref{def-prop:reusable}) it is not intuitive (\ref{def-prop:intuitive}), there is no argument for completeness (\ref{def-prop:complete}), and it is unclear whether it will scale to more complex features (\ref{def-prop:scalable}).
 

### 2018-10-20 fully-abstract-comp/popl2019/paper.tex
Overlay semantics edits

@@
 A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics has on the underlying machine.
-In other words, it should not be possible to have something else than a stack pointer for that part of memory.
+In other words, there should only be stack pointer for the stack memory.
 
@@
 As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
-However, there is no such assumption on the untrusted other components they interact with.
+Untrusted components are not expected to take part in the calling convention.
 To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
-Only instructions at these addresses are allowed to perform calls that push stack frames to the call stack, i.e.\ calls for which \srccm{} provides native LSE and WBCF guarantees.
-The constant $\ta$ is passed around as a parameter of the \srccm{} step relation.
+Only instructions at these addresses can be interpreted as the \srccm{} native call and push frames to the call stack which guarantees LSE and WBCF.
+The constant $\ta$ is a parameter of the \srccm{} step relation.
 Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
@@
   \begin{mathpar}
-    \inferrule{ \src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
+    \inferrule{\src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
                \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
@@
     \hline
-    i \in \Instr                                 & $\sem{i}(\Phi)$ & Conditions\\
+    i \in \src{\Instr}                                 & $\sem{i}(\Phi)$ & Conditions\\
     \hline
@@
     \]
-
     \[
@@
   \caption{Auxiliary definitions used in the operational semantics of \srccm{}.}
-  \label{fig:source-op-sem}
+  \label{fig:source-op-sem-aux}
 \end{figure}

### 2018-10-20 fully-abstract-comp/popl2019/paper.tex
Updated the call return description

@@
 % Call/return
-The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem}, and it may look daunting due to a fair bit of bookkeeping, but it essentially does the following: Given a stack pointer in the stack register, $42$ is pushed to the stack to ensure that the stack frame will be non-empty.
-Based on the stack pointer in $\rstk$, a stack pointer for the callee is created (this corresponds to the return token of \stktokens{}) as well as a return pointer which governs the private part of the stack, i.e.\ addresses $[\aaddr_\stk,\eaddr_\stk]$ which is $42$ and everything (conceptually) below on the stack.
-A new stack frame is pushed to the stack.
-The stack frame consists of the local stack (addresses $[\baddr_\stk,\aaddr_\stk-1]$) as well as the address where execution is supposed to continue on return.
-The return address in the stack frame is an important piece of metadata as it connects the stack frame to the return point and thus allows us on return to check whether we returned to the right place.
-The return pointers are sealed with the seal dictated by $\offpc$ and $\offsigma$ and placed in registers $\rretc$ and $\rretd$, so the callee can use them to return.
-Finally, control is transferred in the same fashion \texttt{xjmp} does it.
-
-The return pair can only be used to return with.
-They only exist on \srccm{} in sealed capabilities, so it is only natural that \texttt{xjmp} takes care of returning.
-To this end, we update the semantics of \texttt{xjmp} by redefining $\xjmpres{}$.
-The function remains the same except when $c_1$ and $c_2$ are code and data return pointers, respectively.
-In this case, the top stack frame is popped from the stack and added to the free stack.
-The execution continues at the return point with a stack pointer in $\rstk$ for the free stack that now includes the memory of the popped stack frame.
-However, the execution only continues if the code return pointer actually points to the correct return address, the base of the returned stack pointer is $\stkb$, and the returned stack pointer is adjacent to the data return pointer (which corresponds to checking the validity of the return token in \stktokens{}).
+The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following: It takes the local stack of the caller and puts it on the call stack along with the return address.
+It also takes the stack capability for the local stack and a copy of the $\pc$-capability adjusted to point to the return address and turns them into a pair of sealed return capabilities.
+The return capabilities are sealed with the seal designated for the call.
+
+The return capabilities, $\retptrc$ and $\retptrd$, can only be used for returning.
+Returns are handled by the \texttt{xjmp} instruction because the call produces sealed return capabilities.
+Upon return, the topmost call stack frame $(\opc,\ms_\var{local})$ is popped from the call stack.
+In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must be in the domain of the local stack.
+If the return succeeds, then the stack pointer is reconstructed, and the local stack is moved back to be part of the active stack.
 
@@
       \src{c_1 = \retptrc(\baddr,\eaddr,\opc)}\\
-      \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge \dom(\ms_\var{local}) \subseteq [\aaddr_\stk,\eaddr_\stk]}\\
+      \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge \dom(\ms_\var{local}) = [\aaddr_\stk,\eaddr_\stk]}\\
       \src{c_\stk = \var{reconstructStackPointer}(\reg(\rstk),c_2) \tand}\\
@@
 \end{multline*}
-  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted).}
+  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted). Auxiliary definitions are found in Figure~\ref{fig:source-op-sem-aux}. }
   \label{fig:source-op-sem}

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Align some equations in Fig. 11.

@@
     \begin{multline*}
-      \src{\ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk = \var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk)} \mathit{\ iff\ } \\
+      \src{\var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk) = \ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk} \mathit{\ iff\ } \\
       \left\{ 
@@
     \end{multline*}
-    \[
-      \src{ \opc, c_\opc = \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} \mathit{\ iff\ }
+    \begin{align*}
+      \src{ \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} &\ \src{= \opc, c_\opc} \mathit{\ iff\ }
       \left\{ 
@@
       \end{array}
-       \right.
-    \]
-    \[
-      \src{\sigma = \var{getCallSeal}(c_\pcreg,\mem,\offpc,\offsigma)} \mathit{\ iff\ }
+       \right.\\
+      \src{\var{getCallSeal}(c_\pcreg,\mem,\offpc,\offsigma)} &\ \src{= \sigma} \mathit{\ iff\ }
       \left\{
@@
       \end{array}
-      \right.
-    \]
-    \[
-     \src{\sealed{\sigma,c_\opc},\sealed{\sigma,c_\var{local}} = \var{sealReturnPair}(\sigma,c_\opc,c_\var{local})}
-    \]
+      \right.\\
+     \src{\var{sealReturnPair}(\sigma,c_\opc,c_\var{local})} &\ \src{= \sealed{\sigma,c_\opc},\sealed{\sigma,c_\var{local}}}
+    \end{align*}%
   % \begin{multline*}
@@
   % \end{multline*}
-
   \begin{multline*}

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Previous commit also put the functions being defined on the left-hand side of equations, rather than on the right.
This commit itself doesn't change anything. I just wanted to add to the log
message of the previous one.
Some minor rephrasings...

@@
 
-The opaque nature of the return pointers is reflected in the interpretation of the instructions common to both \trgcm{} and \srccm{} as \srccm{} does not add special interpretation for them.
+The opaque nature of the return pointers is reflected in the interpretation of the instructions common to both \trgcm{} and \srccm{} as \srccm{} does not add special interpretation for them in non-\texttt{xjmp} instructions.
 Stack pointers, on the other hand, need to behave just like capabilities, so \srccm{} adds new cases for them in the semantics, e.g.\ \texttt{cca} can now also change the current address of a stack pointer as displayed in Figure~\ref{fig:source-op-sem}.
@@
 A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics has on the underlying machine.
-In other words, there should only be stack pointer for the stack memory.
+In other words, there should only be stack pointers for the stack memory.
 
@@
 \srccm{} supports tail calls.
-A tail call is a call from a caller that is done executing and thus doesn't need to be returned to.
+A tail call is a call from a caller that is done executing, and thus doesn't need to be returned to or preserve local state.
 This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in call.
 To perform a tail call, the caller simply transfers control to the callee using \texttt{xjmp}.
-The callee needs to return to the caller's caller, so the caller leaves the return pair they received for the callee to use.
-
-By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from inactive stack frames on the call stack.
-Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack.
-In other words by inspection of the operational semantics, \srccm{} guarantees LSE and WBCF.
-It is important to point this out as \stktokens{} guarantee exactly the LSE and WBCF we have on \srccm{}.
+The tail-callee should return to the caller's caller, so the caller leaves the return pair they received for the callee to use.
+
+It is important to observe that the operational semantics of \srccm{} natively guarantee WBCF (well-bracketed control flow) and (local stack encapsulation) for calls made by trusted components.
+By inspecting the operational semantics of \srccm{}, we can see that it never allow reads or writes to inactive stack frames on the call stack.
+The built-in call for trusted code pushes the local stack frame to the inactive part of the stack, together with the return address. 
+Such frames can be reactivated by \texttt{xjmp}ing to a return capability pair, but only for the topmost stack frame and if the return address corresponds to the one stored in the call stack.
+In other words, WBCF and LSE are natively enforced in this semantics.
 

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Add Acknowledgements.
Line formatting
Clarification of adversary at first use.

@@
 While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
-The caller provides the adversary with a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
+A callee that do not respect the calling convention (hence forward referred to as an adversary\footnote{The assumption on the adversary is described in detail in Section~\ref{sec:well-form-reas}.}) receives a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
@@
 
+\begin{acks}
+  This work was supported by Cost Action CA15123 EUTypes.
+  Dominique Devriese held a Postdoctoral Fellowship from the Research Foundation - Flanders (FWO) during most of this research.
+\end{acks}
+
 \bibliography{references}

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Rewrite seal introduction

@@
 
-Any reasonable capability machine needs a way to set up boundaries between security domains as well as a way to cross the boundaries in a way where each domain retains their authority.
+Any reasonable capability machine needs a way to set up boundaries between security domains with different authorities.
+It also must have a way to cross these boundaries such that (1) the security domain we move from can encapsulate and later regain its authority and (2) the security domain we move to regains all of its authority.
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
 On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
-Syntactically, sealed capabilities add two types of capabilities.
 % seals
-A set of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$ consists of a range of available seals $[\sigma_\var{base},\sigma_\var{end}]$ and the current seal $\sigma$, all represented by a natural number.
-% sealed caps
-A sealed capability $\sealed{\sigma,\vsc}$ consists of a sealable capability $\vsc$ and the seal $\sigma$ it has been sealed with.
-Sealables $\vsc \in \SealableCaps$ are sets of seals or memory capabilities.
+A sealed capability makes an underlying capability opaque which means that the underlying capability cannot be changed or used for the operations it normally gives permission to.
+In other words, the authority the underlying capability represents is encapsulated.
+Syntactically, the sealed capability is represented as $\sealed{\sigma,\vsc}$ where $\vsc$ is a sealable and $\sigma$ is the seal used to seal the capability.
+In order to seal a sealable, it is necessary to have the right to do so.
+The right make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$.
+A set of seal gives the right to seal sealables under seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
+In spirit of the memory capabilities, a set of seals has an active seal $\sigma$ which is the seal currently used to make sealed capabilities.
+We represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
+As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}.
+When capabilities are unsealed with \texttt{xjmp} control will be transferred to a capability that was previously sealed which allows the capability-sealer to decide how the sealed capability is used.
+The \texttt{xjmp} instruction will be explained in more detail later.
+The combination of sealed capabilities and \texttt{xjmp} gives (1) and (2).
 \begin{figure}[tb]

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Rewrite seal introduction

@@
 A sealed capability makes an underlying capability opaque which means that the underlying capability cannot be changed or used for the operations it normally gives permission to.
-In other words, the authority the underlying capability represents is encapsulated.
+In other words, the authority the underlying capability represents is encapsulated under the seal.
 Syntactically, the sealed capability is represented as $\sealed{\sigma,\vsc}$ where $\vsc$ is a sealable and $\sigma$ is the seal used to seal the capability.
 In order to seal a sealable, it is necessary to have the right to do so.
-The right make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$.
+The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$.
 A set of seal gives the right to seal sealables under seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
 In spirit of the memory capabilities, a set of seals has an active seal $\sigma$ which is the seal currently used to make sealed capabilities.
-We represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
+On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
 As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}.
-When capabilities are unsealed with \texttt{xjmp} control will be transferred to a capability that was previously sealed which allows the capability-sealer to decide how the sealed capability is used.
+When underlying capabilities are unsealed with \texttt{xjmp}, control will be transferred to one of the underlying capability which allows the capability-sealer to decide how the sealed capabilities are used.
 The \texttt{xjmp} instruction will be explained in more detail later.

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Add ms to first figure
acknowledgements
commented spurious comment out
minor edit

@@
     \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
-    &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
+\ms \in  &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
     &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
@@
 % KJAA: ^-- Assuming that there is a global stack address capability, no?
-Can I "suddenly" begin to impose \stktokens{}?
+% Can I "suddenly" begin to impose \stktokens{}?
 
@@
 While these are intuitively appealing (\ref{def-prop:intuitive}), it is not clear how they can be used to reason about programs (\ref{def-prop:useful}) or other compiler passes (\ref{def-prop:reusable}), they lack temporal safety guarantees (\ref{def-prop:complete}) and do not scale (\ref{def-prop:scalable}) to settings with dynamic code generation (where a static control flow graph cannot be defined).
-\citet{skorstengaard_reasoning_2017} provide a logical relation that can be used to reason about programs using their calling convention (\ref{def-prop:useful},\ref{def-prop:reusable}) it is not intuitive (\ref{def-prop:intuitive}), there is no argument for completeness (\ref{def-prop:complete}), and it is unclear whether it will scale to more complex features (\ref{def-prop:scalable}).
+\citet{skorstengaard_reasoning_2017} provide a logical relation that
+can be used to reason about programs using their calling convention
+(\ref{def-prop:useful},\ref{def-prop:reusable}), but it is not intuitive (\ref{def-prop:intuitive}), there is no argument for completeness (\ref{def-prop:complete}), and it is unclear whether it will scale to more complex features (\ref{def-prop:scalable}).
 
@@
 \begin{acks}
-  This work was supported by Cost Action CA15123 EUTypes.
+  This research was supported in part by the ModuRes Sapere Aude
+  Advanced Grant from The Danish Council for Independent Research for
+  the Natural Sciences (FNU) and by Cost Action CA15123 EUTypes.
   Dominique Devriese held a Postdoctoral Fellowship from the Research Foundation - Flanders (FWO) during most of this research.

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
comment inserted

@@
 (see below for the meaning of $\Wor \monnefun$).
+\todo[inline]{LB: I don't think it was explained below}
 A region might, for example, require the presence of a certain list of instructions at a given set of memory addresses in both \srccm{} and \trgcm{}.

### 2018-10-21 fully-abstract-comp/popl2019/paper.tex
Resolve a comment by both Lau and Lars.

--- 2018-10-21 fully-abstract-comp/popl2019/paper.tex
@@
 satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$
-% Lau: Where do you return to this? I did not find it very clear what it meant when I was done reading about the worlds.
-(see below for the meaning of $\Wor \monnefun$).
-A region might, for example, require the presence of a certain list of instructions at a given set of memory addresses in both \srccm{} and \trgcm{}.
+(readers unfamiliar with Kripke and step-indexed logical relations may read $\Wor \monnefun \URel{\MemSeg^2}$ as the set of functions from $\Wor$ to relations on memory segments).
+A region might, for example, require the presence of a certain list of instructions at a certain set of memory addresses in both \srccm{} and \trgcm{}.
 If a memory segment is owned by a given region, then that memory segment must be disjoint from the memory segments owned by all other regions.

### 2018-10-22 fully-abstract-comp/popl2019/paper.tex
Proof-read entire paper, making many minor changes along the way.

@@
 \renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
-\renewcommand{\stdreg}[3][]{\iota^{\mathrm{std},#3,#1}_{#2}}
+\renewcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
 \renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
@@
 However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
+
 To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
@@
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not been previously described in the published literature.\footnote{Although they have been mentioned in some technical documents.}
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997} and in technical documents.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 
-A second contribution of this work is the way in which we formulate these two properties: although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
-Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lacked evidence of generality~\cite{skorstengaard_reasoning_2017}.
+A second contribution of this work is the way in which we formulate these two properties.
+Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
+Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
 We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
@@
 The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
-This independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
+Such an independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
 
@@
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015}.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015,watson_fast_2016}.
 % seals
@@
 Syntactically, the sealed capability is represented as $\sealed{\sigma,\vsc}$ where $\vsc$ is a sealable and $\sigma$ is the seal used to seal the capability.
-In order to seal a sealable, it is necessary to have the right to do so.
-The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma}$.
-A set of seal gives the right to seal sealables under seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In spirit of the memory capabilities, a set of seals has an active seal $\sigma$ which is the seal currently used to make sealed capabilities.
-On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
-As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}.
-When underlying capabilities are unsealed with \texttt{xjmp}, control will be transferred to one of the underlying capability which allows the capability-sealer to decide how the sealed capabilities are used.
-The \texttt{xjmp} instruction will be explained in more detail later.
+In order to seal a sealable with a seal $\sigma$, it is necessary to have the right to do so.
+The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
+A set of seals is a capability that represents the authority to seal sealables with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
+In spirit of memory capabilities, a set of seals has an active seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
+% On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
+As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}, an operation that operates on a pair of capabilities sealed with the same seal.
+The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part of the pair) available to the invoked code.
 The combination of sealed capabilities and \texttt{xjmp} gives (1) and (2).
@@
 %%% Domains
-Words on \trgcm{} are capabilities and data (represented by $\ints$).
+Words on \trgcm{} are capabilities and data (represented by integers ($\ints$)).
 We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
 We have complete memories mapping all addresses to words and memory segments mapping some addresses to words (i.e.\ partial functions).
-$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.\ an out-of-bounds memory access.
+$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
 An executable configuration is a register file and memory pair.
@@
 % syntax (instructions)
-\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability-related instructions.
+\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machines as well as capability-related instructions.
 The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
@@
 The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
-The executed instruction is determined by the capability in $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
+The executed instruction is determined by the capability in the $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
 In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
@@
 When a linear capability is moved, the source register must be cleared, so the capability is not duplicated.
-This is done uniformly in the semantics using two approaches.
-In the first one, we use the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
+This is done uniformly in the semantics using the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
 When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
 In the case of \texttt{move}, the source register $\rn$ is overwritten with $\linCons{\Phi(\rn)}$.
-We see the other approach later.
 
@@
 The \texttt{store} and \texttt{load} instructions are fairly standard.
-They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority, and they clear the source of a linear capability transfer.
+They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority.
 Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
@@
 \texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
-It handles linearity by updating the capability in-place, i.e.\ the source register is used as the target register.
-This is the second approach we use to uniformly handle linearity.
-The \texttt{jmp} instruction is a simple jump that just sets the $\pcreg$ register.
+Note that this instruction does not need to use $\linCons{}$ like the previous ones, because it modifies the capability in-place, i.e.\ the source register is also the target register.
+The \texttt{jmp} instruction is a simple jump that just sets register $\pcreg$.
 
@@
   \end{mathpar}
-% \begin{definition}[Initial execution configuration]
-%   \begin{mathpar}
-%     \inference{
-%       c_{\mathrm{main},c} = \sealed{\sigma_1, c_{\mathrm{main},c}'} &
-%       c_{\mathrm{main},d} = \sealed{\sigma_2, c_{\mathrm{main},d}'} &comp\
-%       \sigma_1 = \sigma_2\\
-%       \nonExec{c_{\mathrm{main},d}'}&
-%       \reg(\pcreg) = c_{\mathrm{main},c}' & 
-%       \reg(\rdata) = c_{\mathrm{main},d}' \\
-%       \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
-%       \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
-%       \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0\\
-%       \range{\ms_\stk} = \{0\}&
-%       \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
-%       [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
-%       \overline{\var{import}} = \emptyset
-%     }{
-%       ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
-%     }
-%   \end{mathpar}
-% \end{definition}
 % \begin{definition}[Plugging a program into a context]
@@
 \stktokens{} is based on a traditional single stack, shared between all components.
-As we are on a capability machine, it is natural to add some extra protection to stack and return pointers.
+To explain the technique, let us first consider how we might already add extra protection to stack and return pointers on a capability machine.
 First, we replace stack pointers with stack capabilities.
@@
 While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
-A callee that do not respect the calling convention (hence forward referred to as an adversary\footnote{The assumption on the adversary is described in detail in Section~\ref{sec:well-form-reas}.}) receives a stack capability and a return pair that they are supposed to use for the call, but the adversary can store the provided capabilities on the heap in order to use them later.
+Untrusted callees receive a stack capability and a return pair that they are supposed to use for the call.
+However, a malicious callee (which we will refer to as an adversary\footnote{See Section~\ref{sec:well-form-reas} for more details on our attacker model.}) can store the provided capabilities on the heap in order to use them later.
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
@@
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not seem to provide sufficient guarantees to enforce LSE and WBCF.
+As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
@@
 The linear capabilities are put to use by requiring the stack capability to be linear.
-On call, the caller splits the stack capability in two, such that they have a capability for their local stack frame and a capability for the unused part of the stack.
-The stack capability for their local stack frame is sealed and used as the data part of the sealed return pair, and the stack capability for the remainder of the stack is given to the callee.
+On call, the caller splits the stack capability in two: one capability for their local stack frame and another one for the unused part of the stack.
+The local stack frame capability is sealed and used as the data part of the sealed return pair.
+The capability for the remainder of the stack is given to the callee.
 % Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
@@
 This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
-In the presence of adversarial code, we need some enforcement mechanism that allows us to make sure that the order of the call stack is kept.
+In the presence of adversarial code, we need some mechanism to enforce that the order of the call stack is kept.
 One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
@@
 At this point (Figure~\ref{fig:stack-base-abuse-b}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
+
 For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
@@
 \item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, all return seals are only used for one call site, and they are not leaked.
+\item Reasonable use of seals: Return seals are only used to seal old program pointers, every seal is only used for one call site, and they are not leaked.
 \end{enumerate}
@@
 \label{def-prop:complete}
-\item {\itshape potentially scalable}: even though we currently do not consider dynamic code generation or multi-threading, the current formalization should at least potentially extend to such settings.
+\item {\itshape potentially scalable}: even though we currently do not consider dynamic code generation or multi-threading, the formalization should, at least potentially, extend to such settings.
 \label{def-prop:scalable}
@@
 
-Previous formalisations in the literature are formulated in terms of a static control flow graph~\cite[e.g., ]{Abadi2005Theory}.
+Previous formalisations in the literature are formulated in terms of a static control flow graph~\cite[e.g., ][]{Abadi2005Theory}.
 While these are intuitively appealing (\ref{def-prop:intuitive}), it is not clear how they can be used to reason about programs (\ref{def-prop:useful}) or other compiler passes (\ref{def-prop:reusable}), they lack temporal safety guarantees (\ref{def-prop:complete}) and do not scale (\ref{def-prop:scalable}) to settings with dynamic code generation (where a static control flow graph cannot be defined).
@@
 
-The fact that \stktokens{} enforces LSE and WBCF is formulated as a theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
+The fact that \stktokens{} enforces LSE and WBCF is then formulated as a theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
 The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
@@
 
-Our formal results are complicated by the fact that they only hold for well-behaved components that respect the basic rules of the calling convention, and also on a sane initial configuration of the system.
+Our formal results are complicated by the fact that they only hold on a sane initial configuration of the system and for well-behaved components that respect the basic rules of the calling convention.
 For example, the system should be set up such that seals used by components for constructing return pointers are not shared with other components.
-We envision distributing seals as a job for the linker, so this means our results depend on a linker to do this properly.
+We envision distributing seals as a job for the linker, so this means our results depend on the linker to do this properly.
 As another example, a seal used to construct a return pointer can be reused but only to construct return pointers for the same return point.
@@
 In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components to do certain unsafe things.
+Well-formedness is a requirement for all components (trusted and untrusted), but the reasonability requirement only applies to trusted components, i.e.\ those components for which we provide LSE and WBCF guarantees.
 
@@
 
-\todo[inline]{Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.}
+% see e-mail discussion: we do not see a good way to implement this TODO.
+% \todo[inline]{Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.}
 
@@
 % Describe the things added to the syntax
-The overlay semantics we define for \trgcm{} (\srccm{}) views part of the memory as a built-in stack (Figure~\ref{fig:source-syntax}).
+The overlay semantics \srccm{} for \trgcm{} views part of the memory as a built-in stack (Figure~\ref{fig:source-syntax}).
 % Built-in stack
 To this end, it adds a call stack and a free stack memory to the executable configurations of \trgcm{}.
-The call stack is a list with all the stack frames that are currently inaccessible because they belong to other calls.
-A stack frame contains encapsulated stack memory as well as the program point execution is supposed to return to.
-The free stack memory is the part of the stack that has not been claimed by a call and thus can be used at this point of time.
+The call stack is a list with all the stack frames that are currently inaccessible because they belong to previous calls.
+Every stack frame contains encapsulated stack memory as well as the program point that execution is supposed to return to.
+The free stack memory is the active part of the stack that has not been claimed by a call and thus can be used at this point of time.
 % Stack pointer
@@
 % Call/return
-The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following: It takes the local stack of the caller and puts it on the call stack along with the return address.
-It also takes the stack capability for the local stack and a copy of the $\pc$-capability adjusted to point to the return address and turns them into a pair of sealed return capabilities.
+The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following: It splits the current active stack and the stack capability into the local stack frame of the caller and the rest.
+It also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
+The local stack frame and return address are pushed onto the stack, and turns the local stack capability and return capability into a pair of sealed return capabilities.
 The return capabilities are sealed with the seal designated for the call.
 
-The return capabilities, $\retptrc$ and $\retptrd$, can only be used for returning.
-Returns are handled by the \texttt{xjmp} instruction because the call produces sealed return capabilities.
-Upon return, the topmost call stack frame $(\opc,\ms_\var{local})$ is popped from the call stack.
-In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must be in the domain of the local stack.
+The return capabilities, $\retptrc$ and $\retptrd$ are sealed and can only be used using the \texttt{xjmp} instruction, to perform a return.
+When this happens, the topmost call stack frame $(\opc,\ms_\var{local})$ is popped from the call stack.
+In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must match the domain of the local stack.
 If the return succeeds, then the stack pointer is reconstructed, and the local stack is moved back to be part of the active stack.
@@
        &
-      \srcalt{$\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
+      \srcalt{
          $\begin{multlined}
@@
     $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
+    $\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
     $\sigma =\begin{multlined}[t]
@@
 As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
-The return seals $\sigrets$ are the seals supposed to be used to seal return pointers and the closure seals $\sigcloss$ all the other seals in a component .
-If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjointed from the set of trusted addresses $\ta$, then there should be no return seals.
-The code memory may contain sets of seals but only the seals in $\sigrets$ and $\sigcloss$.
+The return seals $\sigrets$ are the seals supposed to be used to seal return pointers and the closure seals $\sigcloss$ all the other seals in a component.
+If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjoint from the set of trusted addresses $\ta$, then there should be no return seals.
+The code memory may contain sets of seals but only for the seals in $\sigrets$ and $\sigcloss$.
 Other than that, code memory only contains instructions in the form of integers.
@@
 Can adversarial code call into trusted components?
-The answer to that question is yes but not with LSE and WBCF guaranteed.
+The answer to that question is yes but not with LSE and WBCF guarantees.
 Formally, adversarial code can contain the instructions that constitute a call.
@@
 
-We will assume trusted component, for which WBCF and LSE are guaranteed, to be both well-formed and reasonable. 
+We will assume trusted components, for which WBCF and LSE are guaranteed, to be both well-formed and reasonable. 
 %The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.
@@
 
+\begin{figure}
+  \centering
+  \begin{equation*}
+    \inference{
+      c_{\mathrm{main},c},c_{\mathrm{main},d}  = \sealed{\sigma, c_{\mathrm{main},c}', c_{\mathrm{main},d}'} &
+      \nonExec{c_{\mathrm{main},d}'}&
+      \reg(\pcreg,\rdata) = c_{\mathrm{main},c}', c_{\mathrm{main},d}' \\
+      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
+      \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
+      \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0&
+      \range{\ms_\stk} = \{0\}&
+      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
+      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
+      \overline{\var{import}} = \emptyset
+    }{
+      ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
+    }
+  \end{equation*}
+  \caption{The judgement $\var{prog} \rightsquigarrow \Phi$, which defines the initial execution configuration $\Phi$ for executing a program $\var{prog}$.}
+  \label{fig:init-ec}
+\end{figure}
+
 % How a program is lifted to an executable configuration
-Given a program $\var{comp}$, the judgement $\var{comp} \rightsquigarrow \Phi$ defines an initial execution configuration that can be executed.
-It works almost the same on \trgcm{} and \srccm{}.
+Given a program $\var{comp}$, the judgement $\var{comp} \rightsquigarrow \Phi$ in Figure~\ref{fig:init-ec} defines an initial execution configuration that can be executed.
+It works almost the same on \trgcm{} (conditions in \trg{red}) and \srccm{} (conditions in \src{blue}).
 On both machines a stack containing all zeroes is added, as part of the regular memory on \trgcm{} and as the free stack on \srccm{}.
@@
 Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
-\todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}
+% \todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}
 
@@
 To accomodate such conditional relatedness, we construct our relation as a step-indexed Kripke logical relation with recursive worlds.
-Space constraints prevent us from explaining it in full detail in this paper. 
+Space constraints prevent us from explaining this in full detail in this paper. 
 Instead, we will only highlight specific important parts of the logical relation in this section, and we refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
@@
 
-Assumptions about the system, that relatedness is predicated on, are gathered in (Kripke) worlds. 
+Assumptions about the system that relatedness is predicated on, are gathered in (Kripke) worlds. 
 We use a type of worlds that is tailored specifically for our purposes.
@@
 Regions in $\Regionh$ specify the presence of an invariant in the system, shared with the rest of the system.
+They are tagged with the syntactic token $\pure$ and may prescribe two different types of requirements:
 \begin{align*}
@@
                (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
-             \end{multlined}\right.\\
-  \Regions &={}
-             \left\{\begin{multlined}
-               \{\spatial,\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \{\revoked\}
              \end{multlined}\right.
 \end{align*}
-Shared regions are tagged with the syntactic token $\pure$ and may prescribe two different types of requirements.
-First, they may require the presence of \srccm{} and \trgcm{} memory segments
-satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$
-(readers unfamiliar with Kripke and step-indexed logical relations may read $\Wor \monnefun \URel{\MemSeg^2}$ as the set of functions from $\Wor$ to relations on memory segments).
+First, they may require the presence of \srccm{} and \trgcm{} memory segments satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$ (readers unfamiliar with Kripke and step-indexed logical relations may read $\Wor \monnefun \URel{\MemSeg^2}$ as the set of functions from $\Wor$ to relations on memory segments).
 A region might, for example, require the presence of a certain list of instructions at a certain set of memory addresses in both \srccm{} and \trgcm{}.
@@
 Spatial regions are similar to shared regions, but they are tagged as $\spatial$ or $\spatialo$ and may not specify seal invariants.
-Additionally, a spatial region may also be $\revoked$.
+Additionally, a spatial region may also be $\revoked$:
+\begin{align*}
+  \Regions &={}
+             \begin{multlined}
+               \{\spatial,\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \{\revoked\}
+             \end{multlined}
+\end{align*}
+
 The difference between spatial and shared regions is related to linearity and ownership.
 For example, a \trgcm{} linear capability to a piece of memory is related to its \srccm{} counterpart, but only if no other linear capability overlaps with it.
-We will model such an assumption for exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region that only one value in the system may rely on.
-More concretely, we will define how to combine worlds $W_1$ and $W_2$ into a combined world $W_1 \oplus W_2$, on the condition that they represent compatible assumptions: $W_1$ and $W_2$ must contain the same regions except that they must respect exclusive ownership of $\spatialo$ regions: $\pure$ regions must be present in both worlds, but a $\spatialo$ region can only be present in one and must be $\spatial$ in the other.
-We say that worlds $W_1$ and $W_2$ are compatible when $W_1 \oplus W_2$ is defined.
-
-The reason that we have $\spatial$ regions is for defining when \srccm{} and \trgcm{} memories are related.
+We will model such an assumption of exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region that only one value in the system may rely on.
+More concretely, we will define how to combine worlds $W_1$ and $W_2$ into a combined world $W_1 \oplus W_2$, on the condition that they represent compatible assumptions: $W_1$ and $W_2$ must contain the same regions except that they must respect exclusive ownership: $\pure$ regions must be present in both worlds, but a $\spatialo$ region can only be present in one and must be $\spatial$ in the other.
+When $W_1 \oplus W_2$ is defined, we say that worlds $W_1$ and $W_2$ are compatible and we refer to $W_1$ and $W_2$ as compatible partitions of the combined world.
+
+The reason that we have $\spatial$ regions (in addition to $\spatialo$ ones) is for defining when \srccm{} and \trgcm{} memories are related.
 We need them to contain suitable memory segments for all regions in the system, even for regions owned by values that live outside the memory (for example register values).
@@
 
-In our proof, we use only a few regions.
+In our proof, we use only a few different types of regions.
 For space reasons, we do not go into their definitions (see the \citet{technical_report}), but we give a brief overview here.
@@
 The contents of the memory is allowed to change as long as the new contents are still valid.
-Finally, a third type of regions $\stareg[(\ms_S,\ms_T)]{\pure}$ or $\stareg[(\ms_S,\ms_T)]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
+Finally, a third type of regions $\stareg[(\ms_S,\ms_T)]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
 This region is used, for example, to govern inactive parts of the stack whose contents is required to remain unmodified.
@@
 \end{tabular}\\
-These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another (omitted) use of step-indexing.
-For space reasons, we cannot show all of these definitions, but we will try to give a good overview.
-
-Note first how we have two value relations, whose definitions are shown in Figure~\ref{fig:value-relation}.
+These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another use of step-indexing.
+For space reasons, we cannot show all of these definitions, but we will try to give an overview.
+
+Note first how we have two value relations, whose definitions are sketched in Figure~\ref{fig:value-relation}.
 The difference is that the untrusted value relation $\lrv(W)$ does not just express that the two values are related, but also that they are safe to pass to an untrusted adversary, i.e.
 they cannot be used to break LSE and WBCF.
-The trusted value relation does not have the latter requirement and is a strict superset of the former.
-
-Both relations trivially includes numbers $(i,i)$ which are always related to themselves.
+The trusted value relation does not have the latter requirement and is a superset of the former.
+
+Both relations trivially include numbers $(i,i)$ which are always related to themselves.
 The untrusted value relation also includes stack pointers and the underlying linear capability (with the same (non-executable) permission, range of authority, and current address), as well as syntactically equal memory capabilities, seals and sealed values, all under certain conditions involving the world $W$ and the capability's properties.
@@
 Sealed values are in both relations essentially when the sealed values satisfy the relation that was registered for the seal in a region of the world.
-Additionally, when they are combined with any other pair of values related by that relation, the sealed values are required to be related as $\txjmp{}{}$ targets (i.e. in $\lrexj(W)$).
+Additionally, when they are combined with any other pair of values related by that relation, they must be related as $\txjmp{}{}$ targets (i.e. in $\lrexj(W)$).
 Finally, capabilities to code memory are related to themselves in the trusted value relation ($\lrvtrusted(W)$) when there is an appropriate code region in the world.
@@
 In Figure~\ref{fig:obs-rel}, we show sketches of the register file relation $\lrr$, the relation between $\tjmp{}$ targets $\lre(W)$ and the observation relation $\lro$.
-The trusted/untrusted register file relation simply requires that all registers except $\pcreg$ are in the corresponding value relation (in compatible worlds).
+The trusted/untrusted register file relation simply requires that all registers except $\pcreg$ are in the corresponding value relation (in a compatible world partition).
 Two execution configurations are in the observation relation $\lro$ if one terminates whenever the other does\footnote{The actual definition in the \cite{technical_report} is complicated a bit by step-indexing and the fact that we actually use two separate observation relations for left- and right-approximation.}.
@@
   then there exists a $W$ such that
-  $\npair{(\src{\comp},\comp)} \in \lrcomp(W)$
+  $\npair{(\src{\comp},\comp)} \in \lrcomp(W)$.
 \end{lemma}
@@
 
-Finally, we use these two lemmas to prove Theorem~\ref{thm:full-abstraction}
+Finally, we use these two lemmas to prove Theorem~\ref{thm:full-abstraction}.
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
@@
 % - Full abstraction proofs difficult
-Our formulation of WBCF and LSE using a fully abstract overlay semantics has an advantage with respect to others that we haven't discussed yet.
+Our formulation of WBCF and LSE using a fully abstract overlay semantics has an important advantage with respect to others.
 Imagine that you are implementing a fully abstract compiler for a high-level language, i.e.\ a secure compiler that enforces high-level abstractions when interacting with untrusted target-language components.

### 2018-10-22 fully-abstract-comp/popl2019/paper.tex
Address Lau's comments.

@@
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
+As the examples illustrate, this naive use of standard memory and object capabilities does not provide sufficient guarantees to enforce LSE and WBCF.
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
@@
 \item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, every seal is only used for one call site, and they are not leaked.
+\item Reasonable use of seals: Return seals are only used to seal old program pointers, every return seal is only used for one call site, and they are not leaked.
 \end{enumerate}
@@
 It also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
-The local stack frame and return address are pushed onto the stack, and turns the local stack capability and return capability into a pair of sealed return capabilities.
+The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of sealed return capabilities.
 The return capabilities are sealed with the seal designated for the call.
@@
 
-Assumptions about the system that relatedness is predicated on, are gathered in (Kripke) worlds. 
+Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds. 
 We use a type of worlds that is tailored specifically for our purposes.

### 2018-10-22 fully-abstract-comp/popl2019/paper.tex
Add a reference to Szabo's blogpost, small rephrasings.

@@
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997} and in technical documents.
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
-The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see \todo{explain}).
+The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see Section~\ref{sec:well-form-reas}).
 
@@
 Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
-This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
+This includes particularly the line of work on control-flow integrity~\citep{abadi_control-flow_2005}.
 This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.

### 2018-10-22 fully-abstract-comp/popl2019/paper.tex
Rephrase reference to CHERI ISA a bit.

@@
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
+Concurrently with our work, Watson et al.
+have developed a (more realistic) design for linear capabilities in CHERI that will be incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!}.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 
-\todo{Watson et al.\ (XXX: fix name to reflect citation) have concurrently developed an extension of the CHERI-MIPS instruction set with linear capabilities.}
-\todo{\url{https://nakamotoinstitute.org/scarce-objects/}}
-\todo{\url{http://journals.uic.edu/ojs/index.php/fm/article/view/548}}
-
 In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.

### 2018-10-27 fully-abstract-comp/popl2019/paper.tex
Finish final TODO, prepare for submitting 2nd phase review version.

@@
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 \todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
-\todo[inline]{Lau: for camera ready version take a look at the coloring in the paper}
-\todo[inline]{Lau: for camera ready version if TOPLAS paper accepted, update the reference in LR section. }
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
@@
 Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
-Concurrently with our work, Watson et al.
-have developed a (more realistic) design for linear capabilities in CHERI that will be incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!}.
+Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson2018ISA}.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 % Full abstraction thm
-\todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
+% \todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
 \begin{theorem}

### 2018-10-27 fully-abstract-comp/popl2019/paper.tex
Shorten.

@@
 
+\setlength{\belowcaptionskip}{-5pt}
+
 \begin{document}
@@
 
-Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not have been read or modified when the invoked function returns.
+Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not be read or modified until the invoked function returns.
 At the assembly level, this property also does not hold immediately.
@@
 To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
-A first approach uses separate stacks for mutually distrusting components, and a central, trusted stack manager component that mediates cross-component invocations.
+The first is to use separate stacks for mutually distrusting components, and a central, trusted stack manager that mediates cross-component invocations.
 This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it scales poorly to large amounts of components because of the need to reserve separate stack space for all components.
-Also in the presence of higher-order values (e.g.
-function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
-It is not clear how this can be done efficiently in the presence of large amounts of components.
+First, it requires reserving separate stack space for all components, which scales poorly to large amounts of components.
+Also, in the presence of higher-order values (e.g., function pointers, objects), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
+It is not clear how it can do this efficiently in the presence of large amounts of components.
 Finally, this approach does not allow passing stack references between components.
@@
 Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
-Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
+Existing formulations are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
 We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
-Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To accommodate this, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
+It starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
+To this end, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
 This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
@@
 To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
-More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.
-that components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
+More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
 
@@
 Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
-The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
-Such an independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
+The idea is that such a compiler could be proven fully abstract with respect to the well-behaved semantics of the target language, so that the proof could rely on native well-bracketedness and local stack frame encapsulation.
+Such an independent result could then be composed with ours to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
 
@@
   \begin{itemize}
-  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
+  \item \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
   \item proving full-abstraction for the embedding of \srccm{} into \trgcm{} (Section~\ref{sec:fa-proof}) by
@@
 The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
-\begin{wrapfigure}{r}{0.26\linewidth}
+
+\begin{wrapfigure}{r}{0.19\linewidth}
   \centering
@@
   \label{fig:perm-hier}
-\end{wrapfigure}
-
+\end{wrapfigure}%
+%
 %%% Seals
@@
 We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
-We have complete memories mapping all addresses to words and memory segments mapping some addresses to words (i.e.\ partial functions).
+Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ partial functions).
 $\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
@@
 In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
-The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as its range of authority.
+The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities, with $[\var{base},n]$ and $[n+1,\var{end}]$ as ranges of authority.
 Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
@@
   \label{def:program-and-context}
-  We define a \emph{program} to be a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
-
-  A \emph{context} for a component $\comp$ is another component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
+  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
+  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
 \end{definition}
@@
 % write xor exec
-In Section~\ref{sec:form-secur-with}, we will get back to how a program is initialised to create an executable configuration.
+How a program is initialised to create an executable configuration, is discussed in Section~\ref{sec:form-secur-with}.
 % We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
@@
 
-In \stktokens{}, we ensure that at the start of execution the stack memory is only referenced by a single linear stack capability.
-Because of this, we can assert validity of the return token simply by checking its base address and splicing it with the caller's stack frame.
-There is no need to check linearity because we know that only linear capabilities to this memory exist.
+In \stktokens{}, the stack memory is only referenced by a single linear stack capability at the start of execution.
+Because of this, the return token can be verified simply by checking its base address and splicing it with the caller's stack frame.
+There is no need to check linearity because only linear capabilities to this memory exist.
 
@@
 The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
-Both of the capabilities in the pair are of course sealed with the same seal.
+Both of the capabilities in the pair are sealed with the same seal.
 % One seal per return point
-All call points need to be associated with a unique seal (a return seal) that is only used to seal the capabilities in the return pair for that particular call point.
-The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack, which would break WBCF.
+All call points need to be associated with a unique seal (a return seal) that is only used for the return capabilities for that particular call point.
+The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that gave rise to the stack frame, breaking WBCF.
 For similar reasons, we cannot allow return seals to be used to seal closures.
@@
 Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
-This goes for direct leaks, such as leaving a seal in a register or writing it to adversarial memory, as well as indirect leaks, where a capability for reading a return seal (either directly or indirectly) from memory is leaked.
+This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a capability for reading, either directly or indirectly, a return seal from memory).
 
@@
 % Summary of CC
-To summarise, our proposed calling convention \stktokens{} consists of the following measures:
+To summarise, \stktokens{} consists of the following measures:
 \begin{enumerate}
@@
 \end{enumerate}
-Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} with the exception of checking stack base before calls.
-We do not include this check because it only needs to happen once between two calls which means that the check after a call suffices if the base of the stack pointer is not changed subsequently.
+Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} , except for checking stack base before calls.
+We do not include this check because it only needs to happen once between two calls, so that the check after a call suffices if the stack base is not changed subsequently.
 
@@
 \item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
-\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when reasoning about correctness and security of other compiler passes and then compose such results with our results to obtain results about the full compiler.
-\label{def-prop:reusable}
+\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
 \item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
@@
 As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
-Untrusted components are not expected to take part in the calling convention.
+Untrusted components are not assumed to do so.
 To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
@@
 In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must match the domain of the local stack.
-If the return succeeds, then the stack pointer is reconstructed, and the local stack is moved back to be part of the active stack.
+If the return succeeds, the stack pointer is reconstructed, and the local stack becomes part of the active stack again.
 
@@
 Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds. 
-We use a type of worlds that is tailored specifically for our purposes.
+We use a type of worlds tailored to our purposes.
 They consist of three sub-worlds: $\Wor = \Worldh \times \Worlds \times \Worldfs$, capturing assumptions about the heap\footnote{Actually, we use the term heap to describe all memory except the stack, including, for example, code memory.}, the inactive and the active part of the stack, respectively.
@@
 
-The proof of the lemma essentially consists of a big induction where each possible instruction is proven to behave the same in source and target in related memories and register files.
+The proof of the lemma consists of a big induction where each possible instruction is proven to behave the same in source and target in related memories and register files.
 After that first step, the induction hypothesis is used for the rest of the execution.
@@
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  The proof of both directions are very similar, so we only show the right direction.
+  The proofs of both directions are very similar, so we only show the right direction.
   We need to show the \trgcm{} contextual equivalence, so w.l.o.g\ assume a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
@@
 \item a constant and limited amount of checks on every boundary crossing.
-\item possibly a small memory overhead because every stack frame must be of non-zero length
+\item possibly a small memory overhead because stack frames must be of non-zero length
 \end{itemize}
@@
 In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability will need to behave like an atomic operation similar to a compare-and-swap instruction.
-This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
-When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
-If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
+This means that a \texttt{load} of a linear capability should behave atomically, similar to a primitive compare-and-swap instruction.
+This is in principle not a problem except that atomic instructions are significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
+When using \stktokens{}, loads of linear capabilities happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
+Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, well-engineered caching could perhaps reduce the high overhead of atomic loads of linear capabilities.
+% If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
@@
 This includes particularly the line of work on control-flow integrity~\citep{abadi_control-flow_2005}.
-This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
+This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static control graph information and a protected shadow stack.
 Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
@@
 In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
+Also, it only enforces a weak form of control flow correctness saying that jumps stay within the program's static control flow graph~\cite{Abadi2005Theory}.
 Such a property ignores temporal properties and seems hard to use for reasoning.

### 2018-10-27 fully-abstract-comp/popl2019/paper.tex
shorten.

@@
 \label{def-prop:complete}
-\item {\itshape potentially scalable}: even though we currently do not consider dynamic code generation or multi-threading, the formalization should, at least potentially, extend to such settings.
-\label{def-prop:scalable}
+\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, but the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
 \end{enumerate}
@@
 \[
-  \begin{array}{rcl}
-    \offpc,\offsigma & \in & \nats \\
-    \src{\Instr} & \defbnf &  \Instr \mid \scall{\offpc,\offsigma}{r}{r}
+  \begin{array}{rclcrcl}
+    \src{\Instr} & \defbnf &  \Instr \mid \scall{\offpc,\offsigma}{r}{r}&\phantom{skipskipskip}&
+    \offpc,\offsigma & \in & \nats &
   \end{array}
@@
 Using Lemma~\ref{thm:ftlr}, we can now proceed to proving Theorem~\ref{thm:full-abstraction} (full abstraction).
-First, we will extend the logical relation into an omitted relation on components $(\var{comp}_S,\var{comp}_T) \in \lrcomp(W)$.
-Essentially, it relates a component to itself if instantiating its imports with related values produces related exports and code memory satisfying the appropriate code region in the world $W$.
-
+First, we extend the logical relation into an omitted relation on components $(\var{comp}_S,\var{comp}_T) \in \lrcomp(W)$.
+% Essentially, it relates a component to itself if instantiating their imports with related values produces related exports and code memory satisfying the appropriate code region in the world.
 % The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
@@
 % \end{multline*}
-
+%
 Using Lemma~\ref{thm:ftlr} and the definitions of the logical relations, we can then prove the following two lemmas.
@@
 
-Another lemma then relates the component relation and context plugging: given a context and component each related to themselves in the component relation, plugging the context with the component produces an executable configuration in the executable configuration relation.
+Another lemma then relates the component relation and context plugging: plugging related components into related contexts produces related execution configurations.
 \begin{lemma}
@@
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  The proofs of both directions are very similar, so we only show the right direction.
-  We need to show the \trgcm{} contextual equivalence, so w.l.o.g\ assume a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
+  The proofs of both directions are similar, so we only show the right direction.
+  To show the \trgcm{} contextual equivalence, assume w.l.o.g\ a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
   The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.

### 2018-10-27 fully-abstract-comp/popl2019/paper.tex
Shorten.

@@
 To clarify this, we introduce notions of components and programs from which we construct executable configurations.
-A component is basically a program with entry points in the form of imports that need to be linked.
-A component has exports that can satisfy the imports of other components.
-The component definition is in Figure~\ref{fig:target-component-and-linking}.
+A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
+It has exports that can satisfy the imports of other components.
 % Structure
-A base component $\comp_0$ consists of the following: a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses (addresses governed by a linear capability).
+A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses (addresses governed by a linear capability).
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
@@
 %base = library
-A base component has no main entry point, so it needs to be invoked by another component to run. For this reason, it can be seen as a library implementation.
-A component is either a library component or a base component with a main in the form of a pair of sealed capabilities which can be seen as a program that still needs to be linked with libraries.
+A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
+The latter can be seen as a program that still needs to be linked with libraries.
 % Linking .
-Components are combined into new components by linking them together.
-Libraries can be linked together and incomplete programs can be linked with libraries, but two incomplete programs cannot be linked with each other.
+Components are combined into new components by linking them together, as long as only one is an incomplete program with a main.
 Two components can be linked when their memories, seals, and linear addresses are disjoint.
 They are combined by taking the union of each of their constituents.
-For every import that is satisfied by an export of the other component, in the sense that they have the same symbol, the data memory is updated to have the exported word on the imported address.
+For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
 The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
@@
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
-
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for the details.
+Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for details.
 \begin{figure}[tb]
@@
   % Instructions 25-26: pop 42 from the stack.
-  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$.
-  The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
-  The variable $\offsigma$ is the offset in the set of seals to the call seal.
-  The constant $\stkb$ is the globally agreed on stack base.
-  There are a number of ``magic'' numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
-  Line 7: $-5$, offset to the first instruction of the call (from the point the capability was moved out of the $\pcreg$-register).
+  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set of seals it uses and $\offsigma$ the offset in the set of seals to the call seal.
+  $\stkb$ is the globally agreed on stack base.
+  There are some magic numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
+  Line 7: $-5$, offset from line 6 (where $\pcreg$ was copied into $\rtmp{1}$) to line 1.
   Line 12: $5$, offset to the return address.

### 2018-11-03 fully-abstract-comp/popl2019/paper.tex
Title capitalization

--- 2018-10-27 fully-abstract-comp/popl2019/paper.tex
@@
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,showframe]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 
-\setlength{\belowcaptionskip}{-5pt}
-
 \begin{document}
@@
 %% Title information
-\title{\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities}
+\title{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
 % \titlenote{with title note}
@@
 \todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
+\todo[inline]{Lau: for camera ready version take a look at the coloring in the paper}
+\todo[inline]{Lau: for camera ready version if TOPLAS paper accepted, update the reference in LR section. }
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
@@
 
-Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not be read or modified until the invoked function returns.
+Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not have been read or modified when the invoked function returns.
 At the assembly level, this property also does not hold immediately.
@@
 To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
-The first is to use separate stacks for mutually distrusting components, and a central, trusted stack manager that mediates cross-component invocations.
+A first approach uses separate stacks for mutually distrusting components, and a central, trusted stack manager component that mediates cross-component invocations.
 This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it requires reserving separate stack space for all components, which scales poorly to large amounts of components.
-Also, in the presence of higher-order values (e.g., function pointers, objects), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
-It is not clear how it can do this efficiently in the presence of large amounts of components.
+First, it scales poorly to large amounts of components because of the need to reserve separate stack space for all components.
+Also in the presence of higher-order values (e.g.
+function pointers, objects etc.), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
+It is not clear how this can be done efficiently in the presence of large amounts of components.
 Finally, this approach does not allow passing stack references between components.
@@
 In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
-Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson2018ISA}.
+Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although a design sketch for them is incorporated in the latest CHERI ISA reference~\citep{TODO}\todo{reference!} and related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997} and in technical documents.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
-Existing formulations are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
+Formulations in previous work are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
 We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
-It starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To this end, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
+Our new formulation starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
+To accommodate this, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
 This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
@@
 To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
-More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
+More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.
+that components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
 
@@
 Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
-The idea is that such a compiler could be proven fully abstract with respect to the well-behaved semantics of the target language, so that the proof could rely on native well-bracketedness and local stack frame encapsulation.
-Such an independent result could then be composed with ours to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
+The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
+Such an independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
 
@@
   \begin{itemize}
-  \item \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
+  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
   \item proving full-abstraction for the embedding of \srccm{} into \trgcm{} (Section~\ref{sec:fa-proof}) by
@@
 The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
-
-\begin{wrapfigure}{r}{0.19\linewidth}
+\begin{wrapfigure}{r}{0.26\linewidth}
   \centering
@@
   \label{fig:perm-hier}
-\end{wrapfigure}%
-%
+\end{wrapfigure}
+
 %%% Seals
@@
 We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
-Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ partial functions).
+We have complete memories mapping all addresses to words and memory segments mapping some addresses to words (i.e.\ partial functions).
 $\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
@@
 In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
-The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities, with $[\var{base},n]$ and $[n+1,\var{end}]$ as ranges of authority.
+The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities one with $[\var{base},n]$ as its range of authority and the other with $[n+1,\var{end}]$ as its range of authority.
 Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
@@
   \label{def:program-and-context}
-  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
-  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
+  We define a \emph{program} to be a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
+
+  A \emph{context} for a component $\comp$ is another component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
 \end{definition}
@@
 % write xor exec
-How a program is initialised to create an executable configuration, is discussed in Section~\ref{sec:form-secur-with}.
+In Section~\ref{sec:form-secur-with}, we will get back to how a program is initialised to create an executable configuration.
 % We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
@@
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, this naive use of standard memory and object capabilities does not provide sufficient guarantees to enforce LSE and WBCF.
+As the examples illustrate, the standard memory and object-like capabilities (here sealed capabilities) do not provide sufficient guarantees to enforce LSE and WBCF.
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
@@
 
-In \stktokens{}, the stack memory is only referenced by a single linear stack capability at the start of execution.
-Because of this, the return token can be verified simply by checking its base address and splicing it with the caller's stack frame.
-There is no need to check linearity because only linear capabilities to this memory exist.
+In \stktokens{}, we ensure that at the start of execution the stack memory is only referenced by a single linear stack capability.
+Because of this, we can assert validity of the return token simply by checking its base address and splicing it with the caller's stack frame.
+There is no need to check linearity because we know that only linear capabilities to this memory exist.
 
@@
 The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
-Both of the capabilities in the pair are sealed with the same seal.
+Both of the capabilities in the pair are of course sealed with the same seal.
 % One seal per return point
-All call points need to be associated with a unique seal (a return seal) that is only used for the return capabilities for that particular call point.
-The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that gave rise to the stack frame, breaking WBCF.
+All call points need to be associated with a unique seal (a return seal) that is only used to seal the capabilities in the return pair for that particular call point.
+The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack, which would break WBCF.
 For similar reasons, we cannot allow return seals to be used to seal closures.
@@
 Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
-This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a capability for reading, either directly or indirectly, a return seal from memory).
+This goes for direct leaks, such as leaving a seal in a register or writing it to adversarial memory, as well as indirect leaks, where a capability for reading a return seal (either directly or indirectly) from memory is leaked.
 
@@
 % Summary of CC
-To summarise, \stktokens{} consists of the following measures:
+To summarise, our proposed calling convention \stktokens{} consists of the following measures:
 \begin{enumerate}
@@
 \item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, every return seal is only used for one call site, and they are not leaked.
+\item Reasonable use of seals: Return seals are only used to seal old program pointers, every seal is only used for one call site, and they are not leaked.
 \end{enumerate}
-Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} , except for checking stack base before calls.
-We do not include this check because it only needs to happen once between two calls, so that the check after a call suffices if the stack base is not changed subsequently.
+Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} with the exception of checking stack base before calls.
+We do not include this check because it only needs to happen once between two calls which means that the check after a call suffices if the base of the stack pointer is not changed subsequently.
 
@@
 \item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
-\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
+\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when reasoning about correctness and security of other compiler passes and then compose such results with our results to obtain results about the full compiler.
+\label{def-prop:reusable}
 \item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
@@
 As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
-Untrusted components are not assumed to do so.
+Untrusted components are not expected to take part in the calling convention.
 To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
@@
 It also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
-The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of sealed return capabilities.
+The local stack frame and return address are pushed onto the stack, and turns the local stack capability and return capability into a pair of sealed return capabilities.
 The return capabilities are sealed with the seal designated for the call.
@@
 In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must match the domain of the local stack.
-If the return succeeds, the stack pointer is reconstructed, and the local stack becomes part of the active stack again.
+If the return succeeds, then the stack pointer is reconstructed, and the local stack is moved back to be part of the active stack.
 
@@
 % Full abstraction thm
-% \todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
+\todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
 \begin{theorem}
@@
 
-Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds. 
-We use a type of worlds tailored to our purposes.
+Assumptions about the system that relatedness is predicated on, are gathered in (Kripke) worlds. 
+We use a type of worlds that is tailored specifically for our purposes.
 They consist of three sub-worlds: $\Wor = \Worldh \times \Worlds \times \Worldfs$, capturing assumptions about the heap\footnote{Actually, we use the term heap to describe all memory except the stack, including, for example, code memory.}, the inactive and the active part of the stack, respectively.
@@
 Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
-The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see Section~\ref{sec:well-form-reas}).
-
-The proof of the lemma consists of a big induction where each possible instruction is proven to behave the same in source and target in related memories and register files.
+The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see \todo{explain}).
+
+The proof of the lemma essentially consists of a big induction where each possible instruction is proven to behave the same in source and target in related memories and register files.
 After that first step, the induction hypothesis is used for the rest of the execution.
@@
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  The proofs of both directions are very similar, so we only show the right direction.
+  The proof of both directions are very similar, so we only show the right direction.
   We need to show the \trgcm{} contextual equivalence, so w.l.o.g\ assume a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
@@
 \item a constant and limited amount of checks on every boundary crossing.
-\item possibly a small memory overhead because stack frames must be of non-zero length
+\item possibly a small memory overhead because every stack frame must be of non-zero length
 \end{itemize}
@@
 In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability should behave atomically, similar to a primitive compare-and-swap instruction.
-This is in principle not a problem except that atomic instructions are significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
-When using \stktokens{}, loads of linear capabilities happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, well-engineered caching could perhaps reduce the high overhead of atomic loads of linear capabilities.
-% If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
+This means that a \texttt{load} of a linear capability will need to behave like an atomic operation similar to a compare-and-swap instruction.
+This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
+When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
+Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
+If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
 The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
@@
 \section{Related Work}
+
+\todo{Watson et al.\ (XXX: fix name to reflect citation) have concurrently developed an extension of the CHERI-MIPS instruction set with linear capabilities.}
+\todo{\url{https://nakamotoinstitute.org/scarce-objects/}}
+\todo{\url{http://journals.uic.edu/ojs/index.php/fm/article/view/548}}
 
@@
 Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
-This includes particularly the line of work on control-flow integrity~\citep{abadi_control-flow_2005}.
-This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static control graph information and a protected shadow stack.
+This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
+This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
 Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
@@
 In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness saying that jumps stay within the program's static control flow graph~\cite{Abadi2005Theory}.
+Also, it only enforces a weak form of control flow correctness saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
 Such a property ignores temporal properties and seems hard to use for reasoning.

### 2018-11-03 fully-abstract-comp/technical-report.tex

@@
 \title{\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities\\
-  Technical Report with Proves and Details}
+  Technical Report with Proofs and Details}
+\author{Lau Skorstengaard \and Dominique Devriese \and Lars Birkedal}
 \begin{document}
@@
     \tload{\rtmp{1}}{\rtmp{1}}\\
-
     \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
@@
     \txjmp{r_1}{r_2}\\
+
     \text{// the following is the return code}\\
@@
 
+\lau{TODO 18-07-2018: gc should probably be specified here, so they are not just picked at random.}
+\lau{TODO 18-07-2018: Where is this lemma even used?}
 \begin{lemma}[Compatibility lemma for initial execution configuration construction]

### 2018-11-05 fully-abstract-comp/popl2019/paper.tex
Capitalize section headers.

--- 2018-10-27 fully-abstract-comp/popl2019/paper.tex
@@
 
-\subsection{Overlay semantics}
+\subsection{Overlay Semantics}
 % Describe the things added to the syntax
@@
 % \end{figure}
-\subsection{Well-formedness and reasonability}
+\subsection{Well-Formedness and Reasonability}
 \label{sec:well-form-reas}
@@
 
-\subsection{Full abstraction}
+\subsection{Full Abstraction}
 All that is left before we state the full-abstraction theorem is to define how components are combined with contexts and executed, so that we can define contextual equivalence.
@@
 
-\subsection{The logical relation}
+\subsection{The Logical Relation}
 Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
@@
 
-\subsection{Fundamental theorem}
+\subsection{Fundamental Theorem}
 An important lemma in our proof of full abstraction of the embedding of \srccm{} into \trgcm{}, is the fundamental theorem of logical relations (FTLR).
@@
 
-\subsection{Full abstraction proof sketch}
+\subsection{Full Abstraction Proof Sketch}
 \label{subsec:proof-sketch}
@@
 % \end{itemize}
-\subsection{Full abstraction}
+\subsection{Full Abstraction}
 % - Full abstraction proofs difficult
@@
 % ?
-\subsection{Practical applicability}
+\subsection{Practical Applicability}
 We believe there are good arguments for practical applicability of \stktokens{}.

### 2018-11-06 fully-abstract-comp/popl2019/paper.tex
Fix latex error
Prepare the POPL2019 camera-ready version.

@@
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,screen]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %\documentclass[acmsmall]{acmart}\settopmatter{}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002978.10002986.10002989</concept_id>
+<concept_desc>Security and privacy~Formal security models</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10002978.10003006</concept_id>
+<concept_desc>Security and privacy~Systems security</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10003752.10010124.10010138</concept_id>
+<concept_desc>Theory of computation~Program reasoning</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[300]{Security and privacy~Formal security models}
+\ccsdesc[300]{Security and privacy~Systems security}
+\ccsdesc[300]{Theory of computation~Program reasoning}
 
@@
 
+%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
+%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
+%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
+%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
+%%%
+%%% The following is specific to POPL '19 and the paper
+%%% 'StkTokens: Enforcing Well-Bracketed Control Flow and Stack Encapsulation using Linear Capabilities'
+%%% by Lau Skorstengaard, Dominique Devriese, and Lars Birkedal.
+%%%
+\setcopyright{rightsretained}
+\acmYear{2019}
+\copyrightyear{2019}
+\setcopyright{rightsretained}
+\acmJournal{PACMPL}
+\acmYear{2019}
+\acmVolume{3}
+\acmNumber{POPL}
+\acmArticle{19}
+\acmMonth{1}
+\acmPrice{}
+\acmDOI{10.1145/3290332}
+
 \begin{document}
@@
 %% Title information
-\title{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
+\title[\stktokens{}: Enforcing WBCF and LSE using linear capabilities]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
 % \titlenote{with title note}
@@
 
-\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, overlay semantics}
+\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, fully abstract overlay semantics}
 
@@
 %\paragraph{Outline} Blabla
-
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
 
@@
     \src{\Instr} & \defbnf &  \Instr \mid \scall{\offpc,\offsigma}{r}{r}&\phantom{skipskipskip}&
-    \offpc,\offsigma & \in & \nats &
+    \offpc,\offsigma & \in & \nats
   \end{array}
@@
 
+% \begin{acks}                            %% acks environment is optional
+%                                         %% contents suppressed with 'anonymous'
+%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
+%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
+%   %% acknowledge financial support and will be used by metadata
+%   %% extraction tools.
+%   This material is based upon work supported by the
+%   \grantsponsor{GS100000001}{National Science
+%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
+%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
+%   conclusions or recommendations expressed in this material are those
+%   of the author and do not necessarily reflect the views of the
+%   National Science Foundation.
+% \end{acks}
+
 \begin{acks}

### 2018-11-07 fully-abstract-comp/ftlr-proof-cases.tex
typo
Fix the worst typesetting problems in the TR.

@@
   It is also easy to show that if $\Phi_S \sterm[i + \min(j_S,j_T)]{\ta,\stkb}$, then $\Phi_S' \sterm[i]{\ta,\stkb}$ and if $\Phi_T \term[i+\min(j_S,j_T)]$, then $\Phi_T' \term[i]$.
-  The result then follows easily by definition of $\lrol$ and $\lror$.
+  The result then follows easily by definition of $\lro[\preceq,\gc]$ and $\lro[\preceq,\gc]$.
 \end{proof}
@@
 \item $\forall r \in S_{\var{write}} \ldotp \pwheap(r).H \nsupeq
-  \stdreg{R_{\var{write}}(r),\gc}{\pur}.H \wedge \pwheap(r) \text{ is address-stratified}$
+  \stdreg{R_{\var{write}}(r),\gc}{\pur}.H \wedge \pwheap(r) $ is address-stratified
 \end{itemize}
@@
   Additionally, if $\npair{(b,e)}\in\writeCond{l,W}$, then
-  we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_M]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ such that
+  we get an $S'\subseteq \addressable{\lin, \pwheap} \subseteq \activeReg{\pwheap[W_M]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ and 
   for all $r \in S'$, $\pwheap(r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$ and
-  $\pwheap(r) \text{ is address-stratified}$.
+  $\pwheap(r)$ is address-stratified.
 
@@
   $\pwheap[W_M](r').H \nsupeq \stdreg{R'(r'),\gc}{\pur}.H$ and
-  $\pwheap[W_M](r') \text{ is address-stratified}$.
+  $\pwheap[W_M](r') $ is address-stratified.
 
@@
 
-  From this, it follows that $\npair{(\overline{\sigma},\ms_S\update{a}{0},\ms_\var{T,heap}\update{a}{0})} \in \lrheap(W_M)(W_r' \oplus W_\var{heap}')$ and finally $\npair{(\overline{\sigma},\ms_S\update{a}{0},\stk,\ms_\stk,\ms_T\update{a}{0})} \in \lrheap(W_M')(W_r' \oplus W_{\var{heap}}')$.
+  From this, it follows that $\npair{(\overline{\sigma},\ms_S\update{a}{0},\ms_\var{T,heap}\update{a}{0})} \in \lrheap(W_M)(W_r' \oplus W_\var{heap}')$ and finally
+  \begin{equation*}
+    \npair{(\overline{\sigma},\ms_S\update{a}{0},\stk,\ms_\stk,\ms_T\update{a}{0})} \in \lrheap(W_M')(W_r' \oplus W_{\var{heap}}')
+  \end{equation*}
 \end{proof}
@@
   From $\memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}}$, we get an
-  $R_\ms : \dom(\activeReg{\pwfree[W_\var{free\_stack}]}) \fun \MemSeg \times \MemSeg$,
-  $\ms_\var{T,free\_stack} = \biguplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} \pi_2(R_\ms(r))$,
-  $\ms_\stk = \biguplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} \pi_1(R_\ms(r))$,
-  $\stkb \in \dom(\ms_\var{T,free\_stack}) \wedge \stkb \in \dom(\ms_\stk)$,
-  $\exists R_W : \dom(\activeReg{\pwfree[W_\var{free\_stack}]}) \fun \World\ldotp$
-  $W_\var{free\_stack} = \oplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} R_W(r)$,
-  $\forall r \in \activeReg{\pwfree[W_\var{free\_stack}]}$, we have that
+  \begin{itemize}
+  \item $R_\ms : \dom(\activeReg{\pwfree[W_\var{free\_stack}]}) \fun \MemSeg \times \MemSeg$, 
+  \item $\ms_\var{T,free\_stack} = \biguplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} \pi_2(R_\ms(r))$, 
+  \item $\ms_\stk = \biguplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} \pi_1(R_\ms(r))$, 
+  \item $\stkb \in \dom(\ms_\var{T,free\_stack}) \wedge \stkb \in \dom(\ms_\stk)$, 
+  \item $\exists R_W : \dom(\activeReg{\pwfree[W_\var{free\_stack}]}) \fun \World\ldotp$ 
+  \item $W_\var{free\_stack} = \oplus_{r \in \dom(\activeReg{\pwfree[W_\var{free\_stack}]})} R_W(r)$
+  \end{itemize}
+  and for all
+  $r \in \activeReg{\pwfree[W_\var{free\_stack}]}$, we have that
   $\npair[n']{R_\ms(r)} \in  \pwfree[W_\var{free\_stack}](r).H \; \xi^{-1}(R_W(r))$ for all $n' < n$.
@@
   Additionally, if $\npair{(b,e)}\in\stackWriteCond{W}$, then
-  we get an $S'\subseteq \addressable{\lin, \pwfree} \subseteq \activeReg{\pwheap[W_\var{free\_stack}]}$, an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ and
+  we get an 
+  \begin{equation*}
+    S'\subseteq \addressable{\lin, \pwfree} \subseteq \activeReg{\pwheap[W_\var{free\_stack}]} \text{,}
+  \end{equation*}
+  and an $R' : S' \fun \powerset{\nats}$ such that $\biguplus_{r \in S'} R'(r) \supseteq [\baddr,\eaddr]$ and
   for all $r \in S'$, $\pwfree(r).H \nsupeq \stdreg{R'(r),\gc}{\pur}.H$ and
-  $\pwfree(r) \text{ is address-stratified}$.
+  $\pwfree(r)$ is address-stratified.
 
@@
   $\pwfree[W_\var{free\_stack}](r').H \nsupeq \stdreg{R'(r'),\gc}{\pur}.H$ and
-  $\pwfree[W_\var{free\_stack}](r') \text{ is address-stratified}$.
+  $\pwfree[W_\var{free\_stack}](r')$ is address-stratified.
 
@@
   \item case $\isLinear{\ms_S(\aaddr)}$: then $\linCons{\ms_S(\aaddr)} = 0$ and it follows from $\npair[n']{(\ms_S(\aaddr),\ms_T(\aaddr'))}\in\lrv(W)$ that also $\isLinear{\ms_T(\aaddr')}$ and $\linCons{\ms_T(\aaddr')} = 0$.
+
     From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm,\perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrv(W)$, it then follows that $\npair{(\baddr,eaddr)} \in \writeCond{\lin,W}$.
@@
   \begin{itemize}
-  \item case $\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = 0$ and it follows from $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)$ that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
+  \item case $\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = 0$ and it follows from
+    \begin{equation*}
+      \npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)
+    \end{equation*}
+    that also $\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?} and $\linCons{\ms_T(\aaddr')} = 0$.
     From $\linConsPerm{\perm}{\ms_S(\aaddr)}$ and $\linConsPerm{\perm'}{\ms_T(\aaddr')}$, we then also get that $\perm = \perm' \in \writeAllowed{}$ and from $\npair{(c,c')}\in\lrvg{\trust}(H_\sigma,W)$, it then follows that $\npair{(\baddr,eaddr)} \in \stackWriteCond{W}$.
@@
 
-  \item case $\neg\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = \ms_\stk(\aaddr)$ and it follows from $\npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)$ that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?}
+  \item case $\neg\isLinear{\ms_\stk(\aaddr)}$: then $\linCons{\ms_\stk(\aaddr)} = \ms_\stk(\aaddr)$ and it follows from
+    \begin{equation*}
+      \npair[n']{(\ms_\stk(a),\ms_T(a))}\in\lrv(H_\sigma,W)
+    \end{equation*}
+    that also $\neg\isLinear{\ms_T(\aaddr')}$ \dominique{Do we need a lemma for this?}
     and $\linCons{\ms_T(\aaddr')} = \ms_T(\aaddr')$.
@@
         1$
-      \item \label{case:ftlr:move-cca-etc:cca-stack}(restrict,cca,seta2b) $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$,
-        $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, $\Phi_S(r_1) =
-        \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi_T(r_1) =
-        ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and $\perm' \sqsubseteq\perm$,
-        $k = 1$
+      \item \label{case:ftlr:move-cca-etc:cca-stack}(restrict,cca,seta2b)
+        \begin{itemize}
+        \item $w_1 = \stkptr{\perm',\baddr,\eaddr,\aaddr'}$, 
+        \item $w_1' = ((\perm',\linear),\baddr,\eaddr,\aaddr')$, 
+        \item $\Phi_S(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and 
+        \item $\Phi_T(r_1) = ((\perm,\linear),\baddr,\eaddr,\aaddr)$ and 
+        \item $\perm' \sqsubseteq\perm$, 
+        \item $k = 1$
+      \end{itemize}
+
       \item \label{case:ftlr:move-cca-etc:cca-seal}(cca,seta2b) $w_1 = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma'}$,
@@
         = 0$, and $k=3$
-      \item \label{case:ftlr:move-cca-etc:splice-normal}(splice) $\Phi_T(r_2) = \Phi_S(r_2) =
-        ((\perm,\lin),\baddr_2,\eaddr_2,\_)$, and $\Phi_T(r_3) = \Phi_S(r_3) =
+      \item \label{case:ftlr:move-cca-etc:splice-normal}(splice)
+        \begin{itemize}
+        \item $\Phi_T(r_2) = \Phi_S(r_2) = ((\perm,\lin),\baddr_2,\eaddr_2,\_)$,
+          and 
+        \item $\Phi_T(r_3) = \Phi_S(r_3) =
           ((\perm,\lin),\eaddr_2+1,\eaddr_3,\aaddr_3)$, and 
-        $\baddr_2 \leq
-        \eaddr_2$, and $\eaddr_2+1 \leq \eaddr_3$, and
-        $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and $w_2 = w_2' = \linCons{\Phi_S(r_2)}$ and $w_3 = w_3' = \linCons{\Phi_S(r_3)}$, and $k=3$
+        \item $\baddr_2 \leq
+          \eaddr_2$, and 
+        \item $\eaddr_2+1 \leq \eaddr_3$, and
+        \item $w_1=w_1'=((\perm,\lin),\baddr_2,\eaddr_3,\aaddr_3)$, and 
+        \item $w_2 = w_2'
+          = \linCons{\Phi_S(r_2)}$ and 
+        \item $w_3 = w_3' = \linCons{\Phi_S(r_3)}$, and
+        \item $k=3$
+        \end{itemize}
+
       \item \label{case:ftlr:move-cca-etc:splice-seal} (splice) $\Phi_T(r_2) = \Phi_S(r_2) =
@@
       \begin{enumproof}
-      \item(jmp,jnz) $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and
-        $\Phi_T' = \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and
-        $\Phi_S(r_1) = \Phi_T(r_1) = ((\perm_1,\lin_1),\baddr_1,\eaddr_1,\aaddr_1)$,
-        $\exec{\Phi_S(r_1)}$, $\withinBounds{\Phi_S(r_1)}$,
-        $w_1 = \linCons{\Phi_S(r_1)}$ and $w_1' = \linCons{\Phi_T(r_1)}$
-      \item(xjmp) $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) =
-        \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and
-        $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and
-        $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
-        $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$ and
-        $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
-        $\Phi_T'' = \Phi_T\updReg{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}$ and
-        $\Phi_T' = \xjumpResult{c_1'}{c_2'}{\Phi_T''}$
+      \item(jmp,jnz)
+        \begin{itemize}
+        \item $\Phi_S' = \Phi_S\updReg{\pcreg,r_1}{\Phi_S(r_1),w_1}$ and
+        \item $\Phi_T' = \Phi_T\updReg{\pcreg,r_1'}{\Phi_T(r_1),w_1'}$ and
+        \item $\Phi_S(r_1) = \Phi_T(r_1) = ((\perm_1,\lin_1),\baddr_1,\eaddr_1,\aaddr_1)$,
+        \item $\exec{\Phi_S(r_1)}$, 
+        \item $\withinBounds{\Phi_S(r_1)}$,
+        \item $w_1 = \linCons{\Phi_S(r_1)}$ and 
+        \item $w_1' = \linCons{\Phi_T(r_1)}$
+      \end{itemize}
+
+      \item(xjmp)
+        \begin{itemize}
+        \item $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and 
+        \item $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and 
+        \item $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and 
+        \item $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and 
+        \item $c_1' \neq \retptrc(\_)$ and 
+        \item $c_2' \neq \retptrd(\_)$ and 
+        \item $\nonExec{\Phi_S(r_2)}$ and 
+        \item $\nonExec{\Phi_T(r_2)}$ and 
+        \item $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and 
+        \item $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and 
+        \item $\Phi_T'' = \Phi_T\updReg{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}$ and 
+        \item $\Phi_T' = \xjumpResult{c_1'}{c_2'}{\Phi_T''}$
+        \end{itemize}
+
       \end{enumproof}
@@
   \item $\Phi_S.\reg(r) = \sealed{\sigma,\vsc}$ implies that $\Phi_T.\reg(r) = \sealed{\sigma,\vsc'}$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
-  \item $\Phi_S.\reg(r) \neq \retptrd{\_,\_}$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
-  \item $\Phi_S.\reg(r) \neq \retptrc{\_,\_}$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) \neq \retptrd(\_,\_)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
+  \item $\Phi_S.\reg(r) \neq \retptrc(\_,\_,\_)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
   \item $\Phi_S.\reg(r) = \stkptr{\perm_1,\baddr_1,\eaddr_1,\aaddr_1}$ implies that $\Phi_T.\reg(r) = ((\perm_1,\linear),b_1,e_1,a_1)$ and $a \in \dom(\Phi_S.\ms_\stk)$ (follows from $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$).
@@
   In case \ref{case:ftlr:failed}, both executions go to $\failed$.
-  In this case, the result follows vacuously by definition of $\lrol$ and $\lror$.
+  In this case, the result follows vacuously by definition of $\lro[\preceq,\gc]$ and $\lro[\preceq,\gc]$.
   \\\\
@@
 
-  By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r \in S$, $\pwfree[W_{R,1}](r).H \nequal \stdreg{R(r),\gc}{\pur}.H$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
+  By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r \in S$, $\pwfree[W_{R,1}](r).H \nequal \stdreg{R(r),\gc}{\pur}.H$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r)$ is address-stratified.
   Now take $r_\var{priv\_stk}$ fresh and define
@@
 
-          From the $\memSatFStack[n''']{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$ assumption we get $R_\ms : \dom(\activeReg{W_{\var{free\_stack}}''}) \fun \MemSeg \times \MemSeg$ and $R_W : \dom(\activeReg{W_{\var{free\_stack}}''}.\free) \fun \Worlds$ for which
+          From the $\memSatFStack[n''']{\ms^{(3)}_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}''}$ assumption we get
+          \begin{itemize}
+          \item $R_\ms : \dom(\activeReg{W_{\var{free\_stack}}''}) \fun \MemSeg \times \MemSeg$ and
+          \item $R_W : \dom(\activeReg{W_{\var{free\_stack}}''}.\free) \fun \Worlds$
+          \end{itemize}
+          for which
           \begin{itemize}
@@
       \begin{cases}
-        W_{M,\var{free\_stack},[\aaddr_\stk,\eaddr_\stk]}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)]
+        \begin{aligned}
+          W_{M,\var{free\_stack},[\aaddr_\stk,\eaddr_\stk]}&[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked]\\
+          &[\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)]
+        \end{aligned}
+
         & \text{for $r = r_\var{priv\_stk}$} \\
-        R_W(r)[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)] &
+        \begin{aligned}
+          R_W(r)&[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked]\\
+          &[\priv.r_\var{priv\_stk} \mapsto (\stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square},\aaddr+\calllen)]
+        \end{aligned} &
       \end{cases}
@@
 
-  TODO: check that all the below cases take care of the trusted case.
+  \lau{TODO: check that all the below cases take care of the trusted case.}
   \begin{itemize}
@@
         \end{itemize}
-      \item $(\nonLinear{\src{\vsc_S}} \Rightarrow \forall W' \future \purePart{R_R(r_2)}, W_o, n' < n-1, \npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \npair[n']{\vsc_S,\vsc_S',\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))$
+      \item If $\nonLinear{\src{\vsc_S}}$ then for all $W' \future \purePart{R_R(r_2)}$, $W_o$, $n' < n-1$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$ we have that $\npair[n']{\vsc_S,\vsc_S',\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)$
 
         Let $W' \future \purePart{R_R(r_1)}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in \codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \; \sigma \; \xi^{-1}(W_o)$.
-        Further let $n'' \le n'$ be given and assume $\npair[n'']{(\reg_S,\reg_T)} \in \lrrg{\untrusted}(\{\rdata\})(W_R)$, $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
+        Further let $n'' \le n'$ be given and assume that
+        \begin{itemize}
+        \item $\npair[n'']{(\reg_S,\reg_T)} \in
+          \lrrg{\untrusted}(\{\rdata\})(W_R)$
+        \item $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$
+        \end{itemize}
 
@@
           \[
-            \npair[n-1]{((\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk),(\ms_T,\reg_T\update{\pcreg}{\Phi_T(r_1)}\update{\rdata}{\vsc_T'}))} \in \lro
+            \npair[n-1]{
+              \begin{multlined}
+                ((\ms_S,\reg_S\update{\pcreg}{\Phi_S(r_1)}\update{\rdata}{\vsc_S'},\stk,\ms_\stk),\\
+                (\ms_T,\reg_T\update{\pcreg}{\Phi_T(r_1)}\update{\rdata}{\vsc_T'}))
+              \end{multlined}
+            } \in \lro
           \]
@@
     Since $w_2 = w_2'=w_3=w_3'=0$, it's clear that 
-    $\npair{(w_2,w_2') \in \lrvg{\trust}(R_R(r_1))}$ and $\npair{(w_3,w_3') \in \lrvg{\trust}(\purePart{R_R(r_3)})}$.
+    \begin{itemize}
+    \item $\npair{(w_2,w_2') \in \lrvg{\trust}(R_R(r_1))}$ and
+    \item $\npair{(w_3,w_3') \in \lrvg{\trust}(\purePart{R_R(r_3)})}$
+    \end{itemize}
     We take $R_R'(r_1) = R_R(r_2)\oplus R_R(r_3)$, $R_R'(r_2) = R_R(r_1)$ and $R_R'(r_3) = \purePart{R_R(r_3)}$ and $R_R' = R_R(r)$ elsewhere.
@@
 
-    From $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we know that $\npair{(w,w')} \in \lrvg{\trust}(W_{R,2})$, $\npair{(w_2,w_2')} \in \lrvg{\trust}(\purePart{W_{R,2}]})$ and $\npair{(\Phi_S.\reg\update{r_2}{w_2},\Phi_T.\reg\update{r_2}{w_2'})} \in \lrrg{\trust}(W_R')$ with $W_R =W_R' \oplus W_{R,2}$ (using Lemma~\ref{lem:purePart-oplus} and~\ref{lem:purePart-duplicable}).
+    From $\npair{(\Phi_S.\reg,\Phi_T.\reg)} \in \lrrg{\trust}(W_R)$, we know that
+    \begin{itemize}
+    \item $\npair{(w,w')} \in \lrvg{\trust}(W_{R,2})$
+    \item $\npair{(w_2,w_2')} \in \lrvg{\trust}(\purePart{W_{R,2}]})$ and
+    \item $\npair{(\Phi_S.\reg\update{r_2}{w_2},\Phi_T.\reg\update{r_2}{w_2'})} \in \lrrg{\trust}(W_R')$
+    \end{itemize}
+    with $W_R =W_R' \oplus W_{R,2}$ (using Lemma~\ref{lem:purePart-oplus} and~\ref{lem:purePart-duplicable}).
 
@@
       $w_1 = \linCons{\Phi_S(r_1)}$ and $w_1' = \linCons{\Phi_T(r_1)}$
-    \item(xjmp) $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and $\Phi_S(r_2) =
-      \sealed{\sigma,c_2}$ and $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and
-      $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and
-      $c_1' \neq \retptrc(\_)$ and $c_2' \neq \retptrd(\_)$ and
-      $\nonExec{\Phi_S(r_2)}$ and $\nonExec{\Phi_T(r_2)}$ and
-      $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
-      $\Phi_T'' = \Phi_T\updReg{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}$ and
-      $\Phi_T' = \xjumpResult{c_1'}{c_2'}{\Phi_T''}$
+    \item(xjmp)
+      \begin{itemize}
+      \item $\Phi_S(r_1) = \sealed{\sigma,c_1}$ and 
+      \item $\Phi_S(r_2) = \sealed{\sigma,c_2}$ and 
+      \item $\Phi_T(r_1) = \sealed{\sigma,c_1'}$ and
+      \item $\Phi_T(r_2) = \sealed{\sigma,c_2'}$ and
+      \item $c_1' \neq \retptrc(\_)$ and 
+      \item $c_2' \neq \retptrd(\_)$ and
+      \item $\nonExec{\Phi_S(r_2)}$ and 
+      \item $\nonExec{\Phi_T(r_2)}$ and
+      \item $\Phi_S'' = \Phi_S\updReg{r_1,r_2}{\linCons{c_1},\linCons{c_2}}$ and 
+      \item $\Phi_S' = \xjumpResult{c_1}{c_2}{\Phi_S''}$ and
+      \item $\Phi_T'' = \Phi_T\updReg{r_1,r_2}{\linCons{c_1'},\linCons{c_2'}}$ and
+      \item $\Phi_T' = \xjumpResult{c_1'}{c_2'}{\Phi_T''}$
+    \end{itemize}
+
     \end{enumproof}
@@
   \]
-  If $n-1 = 0$, then this holds vacuously (by definition of $\lrol$ and $\lror$), so we assume that $n-1 > 0$.
+  If $n-1 = 0$, then this holds vacuously (by definition of $\lro[\preceq,\gc]$ and $\lro[\preceq,\gc]$), so we assume that $n-1 > 0$.
 

### 2018-11-07 fully-abstract-comp/technical-report.tex
Fix the worst typesetting problems in the TR.

@@
 \maketitle
+
+This document is a technical report accompanying a paper by the same title and authors, published at POPL 2019.
+It contains proofs and details that were omitted from the paper for space and presentation reasons.
+
 \tableofcontents
@@
 \[
-  \lrol = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+  \lro[\preceq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
@@
 \[
-  \lror = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+  \lro[\succeq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
     \begin{array}{l}
@@
     \right\}\cup\\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
     \; \middle| \;
@@
     \right\} \cup \\
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
     \begin{array}{l}
@@
 %
-    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
     \begin{array}{l}
@@
 \[
-  \lrvtrusted[\square,(\ta,\stkb,\gsigrets,\gsigcloss)](W) =
+  \lrvtrusted[\square,\gc](W) =
   \begin{array}[t]{l}
@@
     \begin{array}{l}
+      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
       \exists r \in \dom(\pwheap) \ldotp \\
@@
   By definition of $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$, we get that $W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$ and
-  $\npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
-  \lrheap(\pwheap)(W_{\var{heap}})$ for some $\ms_S' \subseteq \ms_S$ and $\ms_{T,\var{heap}}\subseteq \ms_T$.
+  \begin{equation*}
+    \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in
+    \lrheap(\pwheap)(W_{\var{heap}})
+  \end{equation*}
+  for some $\ms_S' \subseteq \ms_S$ and $\ms_{T,\var{heap}}\subseteq \ms_T$.
 
@@
   \lrheap(\pwheap)(W_{\var{heap}})$, we get an
-  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $\npair[n']{R_\ms(r_{\var{code}})} \in  \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$ for all $n' < n$.
+  $R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg$ and $\exists R_W : \dom(\activeReg{\pwheap}) \fun \World$ such that
+  \begin{itemize}
+  \item $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and
+  \item $\npair[n']{R_\ms(r_{\var{code}})} \in \pwheap(r_{\var{code}}).H \; \xi^{-1}(R_W(r_{\var{code}}))$ for all $n' < n$
+  \end{itemize}
 
@@
   \label{lem:non-linear-pure}
-
+  \hfill
   \begin{itemize}
@@
   \label{lem:conds-shrinkable}
+  \hfill
   \begin{itemize}
@@
     \end{itemize}
-    Since then $W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W'(r)$, we can use Lemma~\ref{lem:oplus-future-distr} to construct an $R_W$ with  $W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)$ and $R_W(r) \future R_W'(r)$ for all $r$.
+    Since then $W = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W'(r)$, we can use Lemma~\ref{lem:oplus-future-distr} to construct an $R_W$ with
+    \begin{equation*}
+      W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r)
+    \end{equation*}
+    and $R_W(r) \future R_W'(r)$ for all $r$.
     Finally, take $R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal}$ to satisfy
@@
   From $\npair{(\var{comp},\var{comp})} \in \lrcomp(W)$, we know that
-  $\{(\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d})\}\subseteq \overline{\var{export}}$ and 
-  $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with $\var{comp}_0 = (\mscode, \msdata, \overline{a_{\mathrm{import}}\mapsfrom s_{\mathrm{import}}}, \overline{s_{\mathrm{export}}\mapsto w_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$.
+  \begin{itemize}
+  \item $\{(\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d})\}\subseteq \overline{\var{export}}$ and
+  \item $\npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W)$ with
+  \item  $\var{comp}_0 = (\mscode, \msdata, \overline{a_{\mathrm{import}}\mapsfrom s_{\mathrm{import}}}, \overline{s_{\mathrm{export}}\mapsto w_{\mathrm{export}}}, \sigrets, \sigcloss,A_\linear)$
+  \end{itemize}
   From this, it follows that for all  $W' \future W$, if
@@
 
-  \item $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$, $\nonExec{\reg_S(\rdata)}$ and $\nonExec{\reg_T(\rdata)}$:
+  \item
+    \begin{itemize}
+    \item $\reg_S(\pcreg) \neq \retptrc(\_)$, 
+    \item $\reg_S(\rdata) \neq \retptrd(\_)$, 
+    \item $\nonExec{\reg_S(\rdata)}$ and 
+    \item $\nonExec{\reg_T(\rdata)}$
+    \end{itemize}
     This follows from the facts that $\reg(\pcreg) = c_{\mathrm{main},c}'$, $\reg(\rdata) = c_{\mathrm{main},d}'$, 
@@
   \item $\npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj[\preceq,\gc](W_\pcreg)$
-  \item $\reg_S(\pcreg) \neq \retptrc(\_)$, $\reg_S(\rdata) \neq \retptrd(\_)$, $\nonExec{\reg_S(\rdata)}$ and $\nonExec{\reg_T(\rdata)}$:
+  \item
+    \begin{itemize}
+    \item $\reg_S(\pcreg) \neq \retptrc(\_)$,
+    \item $\reg_S(\rdata) \neq \retptrd(\_)$, 
+    \item $\nonExec{\reg_S(\rdata)}$ and 
+    \item $\nonExec{\reg_T(\rdata)}$
+    \end{itemize}
   \item $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M}$
@@
     It also follows from the same Lemma~\ref{lem:ftlr-comps} and Lemma~\ref{lem:context-trusted-addr-ampl} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
-    Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$
+    Lemma~\ref{lem:compat-context-plug} then tells us that
+    \begin{equation*}
+      \npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')
+    \end{equation*}
     Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_2} \trg{\term}$, concluding the second direction of the proof.

### 2018-11-08 fully-abstract-comp/popl2019/paper.tex
Further finalization of camera-ready paper.

@@
 %% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,screen]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,screen]{acmart}\settopmatter{}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% use 'none' for review submission.
-\setcopyright{none}
+% \setcopyright{none}
 %\setcopyright{acmcopyright}
@@
 \setcopyright{rightsretained}
-\acmYear{2019}
 \copyrightyear{2019}
-\setcopyright{rightsretained}
 \acmJournal{PACMPL}
@@
 \todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,abate_good_components_2018,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
+Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, \citet{abate_good_components_2018} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
+In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.

### 2018-11-09 fully-abstract-comp/popl2019/paper.tex
Quickly fix some second-phase review comments.

@@
 \label{def-prop:complete}
-\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, but the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
+\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
 \end{enumerate}
@@
 % Call/return
-The interpretation of a call is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following: It splits the current active stack and the stack capability into the local stack frame of the caller and the rest.
-It also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
+The interpretation of $\scall{\offpc,\offsigma}{r_1}{r_2}$ is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following:
+The registers $r_1$ and $r_2$ are expected to contain a code-data pair sealed with the same seal and the unsealed values are invoked by placing them in the $\pcreg$ and $\rdata$ registers, respectively.
+The current active stack and the stack capability are split into the local stack frame of the caller and the rest.
+$\scall{}{}{}$ also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
 The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of sealed return capabilities.
@@
       \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge \dom(\ms_\var{local}) = [\aaddr_\stk,\eaddr_\stk]}\\
-      \src{c_\stk = \var{reconstructStackPointer}(\reg(\rstk),c_2) \tand}\\
+      \src{c_\stk = \var{reconstructStackPointer}(\Phi.\reg(\rstk),c_2) \tand}\\
       \src{c_\opc =  ((\rx,\normal),\baddr,\eaddr,\opc)} 

### 2018-11-20 fully-abstract-comp/popl2019/paper.tex
Update camera-ready version according to editor comments.

@@
 %% Title information
-\title[\stktokens{}: Enforcing WBCF and LSE using linear capabilities]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
+\title[\stktokens{}: Enforcing Well-Bracketed Control Flow and Local State Encapsulation \ldots]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
 % \titlenote{with title note}

### 2018-11-22 fully-abstract-comp/popl2019/paper.tex
Fix title (second try).

--- 2018-11-09 fully-abstract-comp/popl2019/paper.tex
@@
 %% Title information
-\title[\stktokens{}: Enforcing WBCF and LSE using linear capabilities]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
+\title[\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
 % \titlenote{with title note}

### 2019-01-10 presentations/popl19-01-2019/presentation.tex
POPL `19 presentation

\documentclass[aspectratio=169]{beamer}
\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
  }

\usepackage{xcolor}
\input{preamble}

\usepackage{pgfpages}
\setbeameroption{hide notes}
%\setbeameroption{show notes on second screen=right}

\definecolor{myred}{RGB}{192,58,45}
\definecolor{mypurple}{RGB}{180,52,202}
\definecolor{myblue}{RGB}{51,51,179}
\definecolor{mylblue}{RGB}{52,178,204}
\definecolor{mygreen}{RGB}{66,192,103}

\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{calc}
\usepackage{graphicx}

\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape=true,
  literate={\$}{\$}{1},
  keywordstyle=\color{mylblue},
  tabsize=2
}

\lstdefinelanguage
   [x64]{Assembler}     % add a "x64" dialect of Assembler
   [x86masm]{Assembler} % based on the "x86masm" dialect
   % with these extra keywords:
   {morekeywords={CDQE,CQO,CMPSQ,CMPXCHG16B,JRCXZ,LODSQ,MOVSXD, %
                  POPFQ,PUSHFQ,SCASQ,STOSQ,IRETQ,RDTSCP,SWAPGS, %
                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp, %
                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b, %
                  r10,r10d,r10w,r10b,r11,r11d,r11w,r11b, %
                  r12,r12d,r12w,r12b,r13,r13d,r13w,r13b, %
                  r14,r14d,r14w,r14b,r15,r15d,r15w,r15b}} % etc.

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Overview}
    \tableofcontents[currentsection]
  \end{frame}
}

%\input{../illustrations/preamble}
%\input{loc_preamble}
\title{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
\institute{\inst{1} Aarhus University \and %
  \inst{2} Vrije Universiteit Brussel}

\date{POPL, January 19, 2019}

\begin{document}
\maketitle

\newsavebox{\assem}
\begin{lrbox}{\assem}
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
main:
	.cfi_startproc
# BB#0:
	pushq	%rbp
.Ltmp0:
	.cfi\_def\_cfa\_offset 16
.Ltmp1:
	.cfi\_offset %rbp, -16
	movq	%rsp, %rbp
.Ltmp2:
	.cfi\_def\_cfa\_register %rbp
	subq	$\mbox{\textdollar}$16, %rsp
	movabsq	$\mbox{\textdollar}$.L.str, %rdi
	movl	$\mbox{\textdollar}$0, -4(%rbp)
	movb	$\mbox{\textdollar}$0, %al
	callq	printf
	xorl	%ecx, %ecx
	movl	%eax, -8(%rbp)
	movl	%ecx, %eax
	addq	$\mbox{\textdollar}$16, %rsp
	popq	%rbp
	retq
.Lfunc\_end0:
	.size	main, .Lfunc\_end0-main
	.cfi\_endproc
\end{lstlisting}
\end{lrbox}

\newsavebox{\hello}
\begin{lrbox}{\hello}
\begin{lstlisting}[basicstyle=\tiny\ttfamily, language=C, keywordstyle=\color{mylblue}]
#include <stdio.h>
int main()
{
  printf("Hello, World!");
  return 0;
}
\end{lstlisting}
\end{lrbox}

\begin{frame}
  \frametitle{Overview}
  \tableofcontents
\end{frame}

\section{\stktokens{}-paper in the big picture}
\begin{frame}[fragile]
  \frametitle{Abstractions all the way down}
  \centering
  \begin{tikzpicture}
    % \draw (9.5,0.25) rectangle (13.5,4.75);
    \draw (0,0) node {};
    \node<1->[inner sep=0pt, anchor=west] (hardware) at (9.5,2.5) {\includegraphics[angle=90,width=.285\textwidth]{hardware3-2}};

    \draw<2->[dashed] (9,-.5) -- (9,5.5);
%    \draw (4.75,n0.25) rectangle (8.75,4.75);
    \node<2->[inner sep=0pt, anchor=west] (assem) at (4.4,2.5) {\usebox{\assem}};

    \draw<3->[dashed] (3.9,-.5) -- (3.9,5.5);
%    \draw (0,0.25) rectangle (4,4.75);
    \node<3->[inner sep=0pt, anchor=west] (hello) at (0,2.5) {\usebox{\hello}};

    \draw<4> (.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center] node[above] {\phantom{secure}\\compilation} (4.3, 3.7);
    \draw<5-6> (.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center] node[above] {{\color{myred} secure}\\compilation} (4.3, 3.7);
    \node<6->[rectangle,rotate=-25,fill=white,rounded corners,draw] at (11.5,2.5) {{\color{myred}Capability machine}};

    \begin{onlyenv}<7->
      \draw (0.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center,color=white] node[above] {{\color{myred} secure}\\compilation} (4.3, 3.7);
      \draw node[rectangle,fill=white,rounded corners,draw] at (1.93,3.7) (ir1) {ir};
      \draw node[rectangle,fill=white,rounded corners,draw] at (3.15,3.7) (ir2) {ir'};
      \draw (0.8, 3.7) edge[-stealth,bend left,very thick] node[above] {} (ir1);
      \draw (ir1) edge[-stealth,bend left,very thick] node[above] {} (ir2);
      \draw (ir2) edge[-stealth,bend left,very thick] node[above] {} (4.3,3.7);
    \end{onlyenv}

    \draw<8-> node at (4.3,3.7) (p) { };
    \draw<8-> node[rectangle, fit=(ir2) (p),draw={rgb:red,192;green,58;blue,45}] (lastphase) {};
  \end{tikzpicture}

    \note[item]{Compbuter at the hardware level is a lot of physics going on very, very quickly.}
    \note[item]{Abstractions: instead of currents, bits.}
    \note[item]{Assembler - a lot more abstraction. We can make sense of it, but still not nice to work with.}
    \note[item]{High-level languages - enough abstractions to work with it and
      build incredible software systems.}
    \note[item]{In order to move between abstractions, we compile programs}
    \note[item]{Most compilers do not provably preserve the abstractions of
      high-level language.}
    \note[item]{This leads to security vulnerabilities}
    \note[item]{Important for compilation to preserve abstractions.}
    \note[item]{A compilation that preserves abstractions, even when linked with
      unknown assembly,is known as a \emph{secure} compilation}
    \note[item]{Little hope that it is possible to make a secure compilation to
      a traditional processor - they have no way to enforce the abstraction.}
    \note[item]{Necessary with hardware with enforcement mechanisms.}
    \note[item]{Our take is that this could be a capability machine}
    \note[item]{Capability machine explained briefly: Basically a processor as
      you know it except pointers are replaced with capabilities. Capabilities
      do not just point to memory, they also carry permissions that specify what
    operations they can be used for and on what part of the memory they can operate.}
    \note[item]{Naturally, this means that the assembly language must be
      capability aware.}
\end{frame}

% 	.type	.L.str,@object          # @.str
% 	.section	.rodata.str1.1,"aMS",@progbits,1
% .L.str:
% 	.asciz	"Hello, World!"
% 	.size	.L.str, 14

% #include <stdio.h>
% int main()
% {
%   printf("Hello, World!");
%   return 0;
% }

%\section{Paper Overview}
\begin{frame}
  \frametitle{Paper Contents Overview}
  \begin{description}
  \item[Formalization of CHERI-like capability machine with linear capabilities]
  \item[\stktokens{}] a calling convention that provably guarantees local-state
    encapsulation (LSE) and well-bracketed control-flow (WBCF)
  \item[Fully-abstract overlay semantics] a novel way to prove LSE and WBCF claims
  \end{description}
\end{frame}

\section{Defining well-bracketed control flow and local state  encapsulation}

\newcommand{\drawpc}[1]{

}

\begin{frame}[fragile]
  \frametitle{Defining well-bracketed control flow and local state encapsulation}
  \begin{columns}
    \begin{column}{0.5\textwidth}
\begin{lstlisting}[language=C, escapechar=!]
void a()       !\tikz[remember picture] \node[] (bt) {};! 
{
  !\tikz[remember picture] \node[] (pc6) {};!int y = 40+2;
  !\tikz[remember picture] \node[] (pc7) {};!return;      
}              !\tikz[remember picture] \node[] (bb) {};!

void b()
{
  !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
  !\tikz[remember picture] \node[] (pc2) {};!a();
  !\tikz[remember picture] \node[] (pc3) {};!x = 2;
  !\tikz[remember picture] \node[] (pc4) {};!a();
  !\tikz[remember picture] \node[] (pc5) {};!return;   
}
\end{lstlisting}
      \begin{tikzpicture}[remember picture, overlay, text width=3cm]
%        \draw<1> node { };
        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=-1cm] (brace) [midway] {};
        \draw<2> node[right = -0cm of brace] {Function \texttt{a} cannot access variable \texttt{x}};

        \draw<3-> ($(pc1) + (-0.5,0)$) edge[-stealth] node {} (pc1);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.5\textwidth}
      \only<2>{Local-state encapsulation}
      \only<3->{Well-bracketed control flow}
    \end{column}
    % What is LSE and WBCF
    % Quick example?
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Desired properties of the WBCF and LSE definition}
  \begin{enumerate}
\item {\itshape Intuitive}%Should correspond to our intuition about WBCF and LSE
\item {\itshape Useful for reasoning}% we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}.
\item {\itshape Reusable in secure compiler chains}% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
\item {\itshape Arguably "complete"}% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
%\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
\end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Fully-Abstract Overlay Semantics}
  Given
  \begin{itemize}
  \item Target semantics
  \item Calling convention
  \end{itemize}
  Define
  \begin{itemize}
  \item New semantics for the same syntax
  \item Properties by inspection
  \end{itemize}
  In this case:\\
  Given
  \begin{itemize}
  \item Capability machine with linear capabilities
  \item \stktokens{} 
  \end{itemize}
  Define
  \begin{itemize}
  \item Same machine, but with built-in call stack
    \begin{itemize}
    \item Series of instructions corresponding to call and return are
      interpreted as call and return, respectively, using the call stack
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \centering
  Thank you!
\end{frame}

\end{document}

### 2019-01-10 presentations/popl19-01-2019/wbcf/wbcf.c

@@ -1,0 +1,20 @@
+void a()
+{
+  int y = 40+2;
+  return;
+}
+
+void b()
+{
+  int x = 5;
+  a();
+  x = 2;
+  a();
+  return;
+}
+
+int main()
+{
+  a();
+  return 0;
+}

### 2019-01-10 presentations/popl19-01-2019/presentation.tex
Add POPL19 definition evaluation slide

@@
 \begin{frame}
+  \frametitle{Evaluating the definition}
+  \begin{enumerate}
+  \item {\itshape Intuitive}\\
+    \onslide<2->{Yes, call stack corresponds exactly to our intuition.}
+\item {\itshape Useful for reasoning}\\% we should be able to use WBCF and LSE
+                                % when reasoning about correctness and security
+                                % of programs using \stktokens{}.
+  \onslide<3->{Yes, ...}
+\item {\itshape Reusable in secure compiler chains}\\% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
+  \onslide<4->{Yes, fully-abstract compilations compose vertically, so oLCM can
+    be used as a target for other compilation phases.}
+\item {\itshape Arguably "complete"}\\% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
+  \onslide<5->{Yes, ...}
+%\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
+\end{enumerate}
+\end{frame}
+
+\begin{frame}
   \centering

### 2019-01-10 presentations/popl19-01-2019/presentation.tex
Presentation update

@@
 \usepackage{graphicx}
+\usepackage{dashrule}
 
@@
 
-\begin{frame}
-  \frametitle{Fully-Abstract Overlay Semantics}
-  Given
-  \begin{itemize}
-  \item Target semantics
-  \item Calling convention
-  \end{itemize}
-  Define
-  \begin{itemize}
-  \item New semantics for the same syntax
-  \item Properties by inspection
-  \end{itemize}
-  In this case:\\
-  Given
-  \begin{itemize}
-  \item Capability machine with linear capabilities
-  \item \stktokens{} 
-  \end{itemize}
-  Define
-  \begin{itemize}
-  \item Same machine, but with built-in call stack
-    \begin{itemize}
-    \item Series of instructions corresponding to call and return are
-      interpreted as call and return, respectively, using the call stack
-    \end{itemize}
-  \end{itemize}
+\begin{frame}[fragile]
+  \frametitle{Overlay Semantics}
+
+    \begin{tabular}{p{3.2cm} p{3.2cm}}
+\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
+    move  rtmp1 42
+   !\tikz[remember picture] \node[] (calltl) {};!store rstk rtmp1
+    cca   rstk -1
+    geta  rtmp1 rstk
+    cca   rretc 5      !\tikz[remember picture] \node[] (callbr) {};!
+    move  rtmp1 pc
+    cca   rtmp1 -20
+\end{lstlisting}&
+\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
+    load  rtmp1 rtmp1
+    cca   rtmp1 -21
+    cseal rretd rtmp1
+    move  rretc pc
+   !\tikz[remember picture] \node[] (retl) {};!xjmp  r1 r2 !\tikz[remember picture] \node[] (retr) {};!
+    cseal rretc rtmp1
+    move  rtmp1 0
+\end{lstlisting}
+\end{tabular}\tikz[remember picture] \node[] (srcr) {};\\
+\onslide<2->{\hdashrule[0.5ex]{12cm}{1pt}{5pt}}
+\\\begin{tabular}{p{3.2cm} p{3.2cm}}
+\begin{lstlisting}[basicstyle=\tiny\ttfamily]
+    move  rtmp1 42
+    store rstk rtmp1
+    cca   rstk -1
+    geta  rtmp1 rstk
+    cca   rretc 5
+    move  rtmp1 pc
+    cca   rtmp1 -20
+\end{lstlisting}&
+\begin{lstlisting}[basicstyle=\tiny\ttfamily]
+    load  rtmp1 rtmp1
+    cca   rtmp1 -21
+    cseal rretd rtmp1
+    move  rretc pc
+    xjmp  r1 r2
+    cseal rretc rtmp1
+    move  rtmp1 0
+\end{lstlisting}
+\end{tabular}\tikz[remember picture] \node[] (trgr) {};
+    \begin{tikzpicture}[remember picture, overlay]
+      \draw node[rectangle, fit=(calltl) (callbr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (call) {};
+      \draw node[rectangle, fit=(calltl) (callbr)] (calltext) [yshift=-0.1cm] {\textbf{\texttt{call}}};
+      \draw node[right = 3.3cm of trgr, text width = 3cm] {Linear Capability Machine (LCM)};
+
+      \draw node[rectangle, fit=(retl) (retr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (ret) {};
+      \draw node[rectangle, fit=(retl) (retr)] (rettext) [yshift=-0.1cm] {\textbf{\texttt{return}}};
+      \draw node[right = 3.3cm of srcr, text width = 3cm] {Overlay Semantics (oLCM)};
+%      \draw[dashed] (mid) -- ($(mid) + (7,0)$);
+    \end{tikzpicture}
 \end{frame}
@@
 
+% \begin{lstlisting}[basicstyle=\tiny\ttfamily]
+%     move rtmp1 42
+%     store rstk rtmp1
+%     cca rstk -1
+%     geta rtmp1 rstk
+%     split rstk rretd rstk rtmp1
+%     move rtmp1 pc
+%     cca rtmp1 -20
+%     load rtmp1 rtmp1
+%     cca rtmp1 -21
+%     cseal rretd rtmp1
+%     move rretc pc
+%     cca rretc 5
+%     cseal rretc rtmp1
+%     move rtmp1 0
+%     xjmp r1 r2
+%     getb rtmp1 rstk
+%     minus 1 rtmp1 432
+%     move 2 pc
+%     cca rtmp2 5
+%     jnz rtmp2 rtmp1
+%     cca rtmp2 1
+%     jmp rtmp2
+%     fail
+%     splice rstk rdata
+%     cca rstk 1
+%     move rtmp2 0
+% \end{lstlisting}
+
 \end{document}

### 2019-01-10 presentations/popl19-01-2019/presentation.tex
More presentation work

@@
 {
-  !\tikz[remember picture] \node[] (pc6) {};!int y = 40+2;
+  !\tikz[remember picture] \node[] (pc6) {};!...
   !\tikz[remember picture] \node[] (pc7) {};!return;      
@@
   !\tikz[remember picture] \node[] (pc2) {};!a();
-  !\tikz[remember picture] \node[] (pc3) {};!x = 2;
+  !\tikz[remember picture] \node[] (pc3) {};!...
   !\tikz[remember picture] \node[] (pc4) {};!a();
@@
 \begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
-    move  rtmp1 42
+   !\tikz[remember picture] \node[] (srccodetl) {};!move  rtmp1 42
    !\tikz[remember picture] \node[] (calltl) {};!store rstk rtmp1
@@
     cseal rretc rtmp1
-    move  rtmp1 0
+    move  rtmp1 0    !\tikz[remember picture] \node[] (srccodebr) {};!
 \end{lstlisting}
 \end{tabular}\tikz[remember picture] \node[] (srcr) {};\\
-\onslide<2->{\hdashrule[0.5ex]{12cm}{1pt}{5pt}}
+\onslide<2->{\hdashrule[0.1ex]{12cm}{1pt}{5pt}}
 \\\begin{tabular}{p{3.2cm} p{3.2cm}}
@@
     \begin{tikzpicture}[remember picture, overlay]
-      \draw node[rectangle, fit=(calltl) (callbr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (call) {};
-      \draw node[rectangle, fit=(calltl) (callbr)] (calltext) [yshift=-0.1cm] {\textbf{\texttt{call}}};
+      %White overlay
+      \draw<1> node[rectangle,fit=(srccodetl) (srccodebr), fill=white] (whiteoverlay) {};
+
+      % Call red overlay
+      \draw<4-> node[rectangle, fit=(calltl) (callbr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (call) {};
+      \draw<4-> node[rectangle, fit=(calltl) (callbr)] (calltext) [yshift=-0.1cm] {\textbf{\texttt{call}}};
+
+      % LCM text
       \draw node[right = 3.3cm of trgr, text width = 3cm] {Linear Capability Machine (LCM)};
 
-      \draw node[rectangle, fit=(retl) (retr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (ret) {};
-      \draw node[rectangle, fit=(retl) (retr)] (rettext) [yshift=-0.1cm] {\textbf{\texttt{return}}};
-      \draw node[right = 3.3cm of srcr, text width = 3cm] {Overlay Semantics (oLCM)};
+      % Call stack
+      % TODO draw call stack
+      \begin{scope}
+        \clip<3-> ($(srcr) + (0.1,-1.2)$) rectangle ($(srcr) + (2.1,-.75)$);
+        \draw<3-> ($(srcr) + (0.2,-1.5)$) rectangle ($(srcr) + (2,-.75)$) node[pos=.5,align=center] {\scriptsize{...}};
+      \end{scope}
+      \draw<3-> ($(srcr) + (0.2,-0.75)$) rectangle ($(srcr) + (2,-0.25)$) node[pos=.5,align=center] {\scriptsize{frame 1}};
+      \draw<3-> ($(srcr) + (0.2,-0.25)$) rectangle ($(srcr) + (2,0.25)$) node[pos=.5,align=center] {\scriptsize{frame 2}};
+      \draw<3-> ($(srcr) + (0.2,.25)$) rectangle ($(srcr) + (2,.75)$) node[pos=.5,align=center] {\scriptsize{frame 3}};
+
+      \draw<3-> ($(srcr) + (1.1,1.25)$) node[align=center] {\footnotesize{Builtin call stack}};
+      \draw<3-> ($(srcr) + (0.2,-.75)$) -- ($(srcr) + (0.2,-1)$);
+      \draw<3-> ($(srcr) + (2,-.75)$) -- ($(srcr) + (2,-1)$);
+      % Return red overlay
+      \draw<4-> node[rectangle, fit=(retl) (retr),fill={rgb:red,192;green,58;blue,45},fill opacity=0.6, rounded corners] (ret) {};
+      \draw<4-> node[rectangle, fit=(retl) (retr)] (rettext) [yshift=-0.1cm] {\textbf{\texttt{return}}};
+
+      %oLCM text
+      \draw<2-> node[right = 3.3cm of srcr, text width = 3cm] {Overlay Semantics (oLCM)};
 %      \draw[dashed] (mid) -- ($(mid) + (7,0)$);
+    \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Fully-abstract overlay semantics}
+  \centering
+  \begin{tikzpicture}
+    \draw (0,0) node[draw=black, rounded corners] (olcm) {oLCM};
+    \draw node[draw=black, rounded corners,below=2.5cm of olcm] (lcm) {LCM};
+    \draw (olcm) edge[bend left, ->,thick] (lcm);
+    \draw node[left = 3cm of olcm] (ll1) {};
+    \draw node[right = 3cm of olcm] (rl1) {};
+    \draw node[below = 1.25cm of ll1] (ll) {};
+    \draw node[below = 1.25cm of rl1] (rl) {};
+    \draw[dashed] (ll) -- (rl);
+
+    \draw<1> (olcm) edge[bend left, ->,thick] node[right, near start] { } (lcm);
+    \draw<2-> (olcm) edge[bend left, ->,thick] node[right, near start] {$id$} (lcm);
   \end{tikzpicture}
@@
 \end{enumerate}
+\end{frame}
+
+\begin{frame}
+  \frametitle{The \stktokens{} calling convention}
+  \begin{onslide}<2->
+    ... is explained in the paper.
+  \end{onslide}
 \end{frame}

### 2019-01-10 presentations/popl19-01-2019/presentation.tex
Update presentation

@@
 
+\usepackage{appendixnumberbeamer}
 \usepackage{xcolor}
@@
 \usetikzlibrary{calc}
+\usetikzlibrary{arrows,shapes}
 \usepackage{graphicx}
@@
 \begin{lstlisting}[language=C, escapechar=!]
-void a()       !\tikz[remember picture] \node[] (bt) {};! 
+!\tikz[remember picture] \node[] (a) {};!void a()       !\tikz[remember picture] \node[] (bt) {};! 
 {
   !\tikz[remember picture] \node[] (pc6) {};!...
-  !\tikz[remember picture] \node[] (pc7) {};!return;      
+  !\tikz[remember picture] \node[] (pc7) {};!return;!\tikz[remember picture] \node[] (ret) {};!      
 }               !\tikz[remember picture] \node[] (bb) {};!
@@
 \end{lstlisting}
-      \begin{tikzpicture}[remember picture, overlay, text width=3cm]
+      \begin{tikzpicture}[remember picture, overlay]
 %        \draw<1> node { };
-        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=-1cm] (brace) [midway] {};
-        \draw<2> node[right = -0cm of brace] {Function \texttt{a} cannot access variable \texttt{x}};
+        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=0.2cm] (brace) [midway] {};
+        \draw<2> node[right = -0cm of brace, text width = 3cm] {Function \texttt{a} cannot access variable \texttt{x}};
 
-        \draw<3-> ($(pc1) + (-0.5,0)$) edge[-stealth] node {} (pc1);
+        \draw<3> ($(pc4) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc4);
+        \draw<4> ($(a) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (a);
+        \draw<5> ($(pc6) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc6);
+        \draw<6> ($(pc7) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc7);
+        \draw<7-> node[below right = of ret] (w) {\huge ?};
+        \draw<7-> (ret) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202},bend left] (w);
+        \draw<8> ($(pc5) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc5);
+        \draw<9> ($(pc3) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc3);
       \end{tikzpicture}
@@
   \frametitle{Fully-abstract overlay semantics}
-  \centering
+  \begin{columns}
+\begin{column}{0.5\textwidth}
+  \begin{itemize}[<+->]
+  \item Compile from oLCM to LCM
+  \item Compilation is the identity
+  \item Builtin call stack abstraction in oLCM preserved after compilation
+  \end{itemize}
+\end{column}
+\begin{column}{0.5\textwidth}
   \begin{tikzpicture}
@@
   \end{tikzpicture}
+\end{column}
+\end{columns}
 \end{frame}
@@
 
+\appendix
+\begin{frame}
+  \frametitle{Full-abstraction proof sketch}
+  \centering
+    \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
+    % \draw[help lines,yellow] (0,0) grid (10,7);
+    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
+
+    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
+    \node at (5,4) { $\mathrel{\Rightarrow}$ };
+    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
+
+    \node at (4.35,2.8) {  };
+    \node at (5,3.6) {  };
+    \node at (5.65,2.8) {  };
+
+    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
+
+    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
+
+    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
+    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
+    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
+    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
+    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
+
+    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
+    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
+    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
+
+    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
+
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em, below=0.5cm]{\small Contextual equivalence preservation} (0,1);
+  \end{tikzpicture}
+\end{frame}
+
 \end{document}

### 2019-01-12 presentations/illustrations/linear-prevention-attack-1.tex
New version of POPL19 presentation

--- 2017-12-14 presentations/illustrations/linear-prevention.tex
@@
 \begin{frame}
-  \frametitle{Linear caps: stack pointers}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[fill=gray!50, right=0cm of sp2] { return data pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actsf{(0,5)}{(6,6.5)}{callee stack frame}
-    \end{onlyenv}
-
-    % Return from callee
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Caller splices the stack frames
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-    \end{onlyenv}
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-    \frametitle{Linear caps prevent attack 1}
+    \frametitle{Linear capabilities prevent the attack 1}
   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
 \end{frame}
-
-\begin{frame}
-  \frametitle{Linear caps prevents attack 2}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    \draw (16.5,5.5) node {register file};
-    \begin{scope}
-      \clip (15.9,1.1) rectangle (17.1,5.1);
-      \foreach \x in {1,2,...,4}
-      {
-        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
-      };
-    \end{scope}
-    \node (r1) at (16.75,4.5) {};
-    \node (r2) at (16.75,3.6) {};
-
-    % Caller 
-    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
-    \begin{onlyenv}<+>
-      \capbrace{(6.5,2.5)}{(6.5,16)}
-      \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,thick,bend left] (sp1);
-    \end{onlyenv}
-
-    % Split the stack pointer
-    \begin{onlyenv}<+>
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw (r2) edge[->,red,thick,bend left] (sp2);
-      \draw (r1) edge[->,red,thick,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \draw node[right=0cm of sp2] { private stack pointer };
-    \end{onlyenv}
-
-    % Callee called
-    \begin{onlyenv}<+>
-      \capbrace[usp]{(6.5,5)}{(6.5,16)}
-      \draw (r1) edge[->,red,bend left] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    \begin{onlyenv}<.->
-      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray!50,right=0cm of sp2] { \phantom{return data pointer} };
-      \draw (r2) edge[->,red,bend left] (sp2);
-      \draw node[right=0cm of sp2] { return data pointer };
-    \end{onlyenv}
-
-    \begin{onlyenv}<+>
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \actadv{(0,5)}{(6,8)}{callee stack frame}
-    \end{onlyenv}
-
-    % At this point we have a linear point stored in a part of memory that it governs.
-In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
-%    \draw<+> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
-
-    \begin{onlyenv}<+->
-
-      \capbrace[usp]{(13.5,5)}{(13.5,16)}
-      \draw (6,13.25) edge[red,thick,->,in=45] (usp);
-      \draw node[right=0cm of usp] { stack pointer };
-
-      \begin{scope}
-        \fill[fill=gray!50] (0,2.5) rectangle (6,15) node[pos=.5,color=black] {};
-        \draw (0,0) -- (0,15);
-        \draw (6,0) -- (6,15);
-      \end{scope}
-      \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize sp};
-      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
-      \inactadv{(0,5)}{(6,8)}{callee stack frame}
-      \actsf{(0,8)}{(6,10.5)}{callee (2) stack frame} 
-  \end{onlyenv}
-
-  \draw<+>[teal,very thick] (-0.5,7.5) -- (6.5,11);
-  \draw<.>[teal,very thick] (-0.5,11) -- (6.5,7.5);
-  \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { No stack pointer\\for callee (2)!};
-
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Linear caps prevents attack 3}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-  \scope
-    \clip (-.1,-.1) rectangle (6.1,15.1);
-    \fill[fill=white] (0,0) rectangle (6,15);
-    \draw (0,0) -- (0,15);
-    \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,1.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
-  \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-
-  \begin{scope}
-    \fill[gray!50] (13,0) rectangle (18,13);
-    \draw (13,0) -- (13,13);
-    \draw (18,0) -- (18,13);
-    \draw (15.5,14) node {heap memory};
-  \end{scope}
-
-\note{Caller calls callee (split stack pointer to make ret-ptr-d)}
-  \begin{onlyenv}<+->
-    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
-    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-  \end{onlyenv}
-  \begin{onlyenv}<.>
-    \capbrace[sp]{(6.1,1.5)}{(6.1,15)}
-    \draw node[right=0cm of sp] { sp };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \capbrace[ret1]{(6.1,1.5)}{(6.1,4.5)}
-    \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr-d} };
-    \draw node[right=0cm of ret1] { ret-ptr-d };
-  \end{onlyenv}
-  \begin{onlyenv}<.->
-    \draw [decorate,decoration={brace,aspect=0.3,amplitude=4pt,raise=4pt},yshift=0pt]
-    (12.9,1) -- (12.9,3) node[draw=black] (ret1c) [black,pos=0.3,xshift=-0.6cm] {};
-    \draw node[fill=gray!50,left=0cm of ret1c] { \phantom{ret-ptr-c} };
-    \draw node[left=0cm of ret1c] { ret-ptr-c };
-  \end{onlyenv}
-
-  \begin{onlyenv}<.-.(3)>
-    \capbrace[sp]{(6.1,4.5)}{(6.1,15)}
-    \draw node[right=0cm of sp] { sp };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-    \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
-    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
-  \end{onlyenv}
-
-\note{Callee saves ret-ptr-d on the heap.}
-  \begin{onlyenv}<+->
-    \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
-  \end{onlyenv}
-\note{Callee saves ret-ptr-c on the heap.}
-  \begin{onlyenv}<+->
-    \draw (13,4.75) edge[red,bend right,thick,->] (ret1c);
-  \end{onlyenv}
-
-\note{Callee calls callee (1)}
-\note{Before the callee jumps, it is assumed that it has checked the global "max height" of its stack pointer!}
-  \begin{onlyenv}<+->
-    \capbrace[ret2]{(6.1,4.5)}{(6.1,7.5)}
-    \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr-d (1)} };
-    \draw node[right=0cm of ret2] { ret-ptr-d (1) };
-  \end{onlyenv}
-  \begin{onlyenv}<.-.(1)>
-    \capbrace[sp]{(6.1,7.5)}{(6.1,15)}
-    \draw node[right=0cm of sp] { sp };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
-    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
-    \actsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
-    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-    \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-  \end{onlyenv}
-
-\note{Callee (1) calls callee (2)}
-  \begin{onlyenv}<+->
-    \capbrace[ret3]{(6.1,7.5)}{(6.1,10.5)}
-    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
-    \draw node[right=0cm of ret3] { ret-ptr (2) };
-  \end{onlyenv}
-  \begin{onlyenv}<.->
-    \capbrace[sp]{(6.1,10.5)}{(6.1,15)}
-    \draw node[right=0cm of sp] { sp };
-  \end{onlyenv}
-
-  \begin{onlyenv}<+->
-    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
-    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
-    \inactsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
-
-    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
-    \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
-    \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
-  \end{onlyenv}
-  \begin{onlyenv}<.-.(1)>
-    \actadv{(0,10.5)}{(6,13.5)}{callee stack frame (2)}
-  \end{onlyenv}
-
-\note{Callee (2) has a capability for Callee and can thus access the stored ret-ptr-d}
-  \begin{onlyenv}<+->
-    \draw [decorate,decoration={brace,amplitude=4pt,mirror,raise=4pt},yshift=0pt]
-    (18.1,4) -- (18.1,6) node[draw=black] (calleecap) [black,midway,xshift=0.6cm] {};
-    \draw (18,11) edge[red,bend left,thick,->] (calleecap);
-  \end{onlyenv}
-
-\note{Callee (2) uses ret-ptr-d to return}
-  \begin{onlyenv}<+->
-    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
-    \draw[fill=gray!30] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
-    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
-    \draw node[fill=white,right=0cm of ret1] { ret-ptr-d };
-    \draw node[fill=white,left=0cm of ret1c] { ret-ptr-c };
-    \draw (13,4.75) edge[red,bend right,thick,->] (ret1);
-  \end{onlyenv}
-\note{Splicing: fails because ends of capabilities don't meet. callee stack frame (1) cannot be empty (it will at least contain 42).}
-  \begin{onlyenv}<+->
-    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (8,7.5) {Caller tries to splice sp \\and ret-ptr-d, but fails!};  
-  \end{onlyenv}
-\note{Before the caller makes another call/return, it will check that the stack height matches the global max stack height.}
-\end{tikzpicture}
-
-\end{frame}
-
-%%% Local Variables:
-%%% TeX-master: "presentation-illustrations"
-%%% End:

### 2019-01-12 presentations/popl19-01-2019/presentation.tex
New version of POPL19 presentation

@@
 \input{preamble}
+\input{../illustrations/preamble}
 
@@
       \draw (0.8, 3.7) edge[-stealth,out=25,in=155,very thick,align=center,color=white] node[above] {{\color{myred} secure}\\compilation} (4.3, 3.7);
-      \draw node[rectangle,fill=white,rounded corners,draw] at (1.93,3.7) (ir1) {ir};
-      \draw node[rectangle,fill=white,rounded corners,draw] at (3.15,3.7) (ir2) {ir'};
+      \draw node[rectangle,fill=white,draw] at (1.93,3.7) (ir1) {...};
+      \draw node[rectangle,fill=white,rounded corners,draw] at (3.15,3.7) (ir2) {ir};
       \draw (0.8, 3.7) edge[-stealth,bend left,very thick] node[above] {} (ir1);
@@
 
+% Secure compilations: If security properties in target can be proven in source.
+
 % 	.type	.L.str,@object          # @.str
@@
 %\section{Paper Overview}
+% \begin{frame}
+%   \frametitle{Linear capability machine}
+%   \begin{columns}
+% \begin{column}{0.5\textwidth}
+%   \begin{itemize}
+%   \item<1-> Low-level machine
+%     \note<1->[item]{Capability machines are low-level machines proposed in the systems community.}
+%     \note<1->[item]{For instance, the CHERI OS operates on one.}
+%     \note<1->[item]{Has all the instructions we expect, load, store, jmp, etc.}
+%     % \note{Has instructions as you would expect, load, store, jmp, etc.}
+%   \item<2-> Capabilities replace pointers
+%     \begin{itemize}
+%     \item<3-> Pointer
+%     \item<4-> Range of authority
+%     \item<5-> Kind of authority
+%       \note<5->[item]{Roughly two kinds of capabilities:}
+%       \begin{itemize}
+%       \item read, write, and execute
+%         \note<5->[item]{Memory capabilities, allows you to do all the standard memory operations.}
+%       \item sealed
+%         \note<5->[item]{Provides encapsulation mechanism which allows seperation of security domains.}
+%         \note<5->[item]{Can not be used for anything but jump, when jumped to becomes read/execute.}
+%       \end{itemize}
+%     \end{itemize}
+%     \item<6-> Capability manipulation instructions
+%     \item<7-> Authority checked dynamically
+%     \item<8-> Linear capabilities non-copyable
+%       \begin{itemize}
+%       \item<9-> Split and splice instructions
+%       \end{itemize}
+%     \end{itemize}
+% \end{column}
+
+% \begin{column}{0.5\textwidth}
+%   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+%     \draw (1,9) node {Memory};
+%     \draw (-3.5,0) node {\phantom{.}};
+%     \scope
+%       \clip (-.1,-.1) rectangle (6.1,8.1);
+%       \fill[fill=white] (0,0) rectangle (6,8);
+%       \draw (0,0) -- (0,8);
+%       \draw (6,0) -- (6,8);
+%     \endscope
+%     % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+%     \draw (-1.5,0) node {};
+%     \draw (-1.5,8) node {};
+%     \foreach \x in {0.5,1,...,7}
+%     {
+%       \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
+%     };
+
+%     \draw<3>  (-3,1.5) edge[black,->,in=180] (0,2.75);
+
+%     \draw<4-9> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
+%     (-0.25,1) -- (-0.25,4.5) node[draw=black] (sp1) [black,midway,xshift=-12pt] {};
+%     \draw<4>  (-3,1.5) edge[black,->,in=180] node[above left]
+%     {\footnotesize\phantom{\textsc{rw}}} (sp1);
+%     \draw<5-7>  (-3,1.5) edge[black,->,in=180] node[above left] {\footnotesize \textsc{rw}} (sp1);
+
+%     \draw<8-9>  (-3,1.5) edge[black,->,in=180, dashed] node[above left] {\footnotesize \textsc{rw}} (sp1);
+
+%     \draw<10> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
+%     (-0.25,1) -- (-0.25,2.5) node[draw=black] (sp2) [black,midway,xshift=-12pt] {};
+%     \draw<10> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
+%     (-0.25,2.5) -- (-0.25,4.5) node[draw=black] (sp3) [black,midway,xshift=-12pt] {};
+%     \draw<10>  (-3,2) edge[black,->,in=180, dashed] node[above left] {\footnotesize \textsc{rw}} (sp3);
+%     \draw<10>  (-3,1) edge[black,->,in=180, dashed] node[below right] {\footnotesize \textsc{rw}} (sp2);
+
+%     \draw<11-> [decorate,decoration={brace,amplitude=3pt,raise=0pt},yshift=0pt]
+%     (-0.25,1) -- (-0.25,4.5) node[draw=black] (sp1) [black,midway,xshift=-12pt] {};
+%     \draw<11->  (-3,1.5) edge[black,->,in=180, dashed] node[above left] {\footnotesize \textsc{rw}} (sp1);
+%   \end{tikzpicture}
+% \end{column}
+% \end{columns}
+% \end{frame}
+\section{Capability machines and \stktokens{}}
+
+\input{../illustrations/traditional}
+\input{../illustrations/capabilities}
 \begin{frame}
-  \frametitle{Paper Contents Overview}
-  \begin{description}
-  \item[Formalization of CHERI-like capability machine with linear capabilities]
-  \item[\stktokens{}] a calling convention that provably guarantees local-state
-    encapsulation (LSE) and well-bracketed control-flow (WBCF)
-  \item[Fully-abstract overlay semantics] a novel way to prove LSE and WBCF claims
-  \end{description}
+    \frametitle{\stktokens{} prevent the attack}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+    \draw (16.5,5.5) node {register file};
+    \begin{scope}
+      \clip (15.9,1.1) rectangle (17.1,5.1);
+      \foreach \x in {1,2,...,4}
+      {
+        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
+      };
+    \end{scope}
+    \node (r1) at (16.75,4.5) {};
+    \node (r2) at (16.75,3.6) {};
+
+    % Caller 
+    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+    \begin{onlyenv}<+>
+      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (r1) edge[->,red,thick,bend left] (sp1);
+    \end{onlyenv}
+
+    % Split the stack pointer
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw (r2) edge[->,red,thick,bend left] (sp2);
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw (r1) edge[->,red,thick,bend left] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    % Callee called
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw (r1) edge[->,red,bend left] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
+      \draw node[right=0cm of sp2] { return data pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+    \draw<.-> (r2) edge[->,red,bend left] (sp2);
+
+    \begin{onlyenv}<.->
+      \begin{scope}
+        \fill[gray!20] (14,9) rectangle (18,14);
+        \draw (14,9) -- (14,14);
+        \draw (18,9) -- (18,14);
+      \end{scope}
+    \end{onlyenv}
+    \draw<.>[fill=white] (14,10) rectangle node[color=red] {} (18,10.5);
+
+    % Callee stores stack pointer
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (8,5) -- (8,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw (16,15) node {heap memory};
+    \end{onlyenv}
+
+    \begin{onlyenv}<.>
+      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
+      \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
+      \draw node[right=0cm of usp] { stack pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
+      \draw node[right=0cm of sp2] { return data pointer };
+    \end{onlyenv}
+
+    % Return from callee
+    \begin{onlyenv}<+->
+      \draw (14,10.25) edge[->,red,bend right] (usp);
+      \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
+      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[right=0cm of sp2] { private stack pointer };
+      \draw node[right=0cm of usp] { stack pointer };
+    \end{onlyenv}
+
+    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
+    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
+    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
+  \end{tikzpicture}  
 \end{frame}
 
-\section{Defining well-bracketed control flow and local state  encapsulation}
+% \begin{frame}
+%   \frametitle{\stktokens{}}
+%  \begin{itemize}
+%  \item Use linear capability for stack.
+% \item On call, split off local stack frame and encapsulate it.
+% \item   On return, require that the local stack frame is returned.
+% \item   The encapsulated stack frame works like a token that permits you to return. You must use the token when returning.
+% \item   A good number of details to get right - see the paper.
+%    \end{itemize}
+% \end{frame}
 
-\newcommand{\drawpc}[1]{
-
-}
-
+\section{Defining well-bracketed control flow and local state encapsulation by fully-abstract overlay semantics}
 \begin{frame}[fragile]
-  \frametitle{Defining well-bracketed control flow and local state encapsulation}
+  \frametitle{Well-bracketed control flow and local state encapsulation}
   \begin{columns}
@@
 \end{lstlisting}
+% Mention that our work support high-order
+% No good definition. We appeal to you intuitive definition
       \begin{tikzpicture}[remember picture, overlay]
@@
     \begin{column}{0.5\textwidth}
-      \only<2>{Local-state encapsulation}
-      \only<3->{Well-bracketed control flow}
+      \only<2>{Local-state encapsulation (LSE)}
+      \only<3->{Well-bracketed control flow (WBCF)}
     \end{column}
@@
 
-\begin{frame}
-  \frametitle{Desired properties of the WBCF and LSE definition}
-  \begin{enumerate}
-\item {\itshape Intuitive}%Should correspond to our intuition about WBCF and LSE
-\item {\itshape Useful for reasoning}% we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}.
-\item {\itshape Reusable in secure compiler chains}% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
-\item {\itshape Arguably "complete"}% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
-%\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
-\end{enumerate}
-\end{frame}
-
 \begin{frame}[fragile]
-  \frametitle{Overlay Semantics}
-
+  \frametitle{Fully-abstract overlay Semantics}
+\onslide<8->{\hdashrule[0.1ex]{12cm}{1pt}{5pt}}
 \begin{tabular}{p{3.2cm} p{3.2cm}}
@@
     cca   rtmp1 -20
-\end{lstlisting}&
+\end{lstlisting}& \tikz[remember picture] \node[] (srcm) {};
 \begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
@@
     cca   rtmp1 -20
-\end{lstlisting}&
+\end{lstlisting}& \tikz[remember picture] \node[] (trgm) {};
 \begin{lstlisting}[basicstyle=\tiny\ttfamily]
@@
       % LCM text
-      \draw node[right = 3.3cm of trgr, text width = 3cm] {Linear Capability Machine (LCM)};
+      \draw node[right = 3.3cm of trgr, text width = 3cm] {Linear Capability Machine};
 
       % Call stack
-      % TODO draw call stack
       \begin{scope}
@@
       \draw<3-> ($(srcr) + (2,-.75)$) -- ($(srcr) + (2,-1)$);
+
       % Return red overlay
@@
       %oLCM text
-      \draw<2-> node[right = 3.3cm of srcr, text width = 3cm] {Overlay Semantics (oLCM)};
+      \draw<2-> node[right = 3.3cm of srcr, text width = 3cm] {Overlay Semantics};
       % \draw[dashed] (mid) -- ($(mid) + (7,0)$);
+
+      % Compilation, no id
+      \draw<5> ($(srcm) + (-0.3,-1)$) edge[bend left, -stealth, very thick] ($(trgm) + (-0.3,-1)$);
+
+      % Compilation with id
+      \draw<6-> ($(srcm) + (-0.3,-1)$) edge[bend left, -stealth, very thick] node[right] {$id$} ($(trgm) + (-0.3,-1)$);
+
+      % Call stack
+      \begin{scope}
+        \clip<7-> ($(trgr) + (0.1,-1.2)$) rectangle ($(trgr) + (2.1,-.75)$);
+        \draw<7-> ($(trgr) + (0.2,-1.5)$) rectangle ($(trgr) + (2,-.75)$) node[pos=.5,align=center] {\scriptsize{...}};
+      \end{scope}
+      \draw<7-> ($(trgr) + (0.2,-0.75)$) rectangle ($(trgr) + (2,-0.25)$) node[pos=.5,align=center] {\scriptsize{frame 1}};
+      \draw<7-> ($(trgr) + (0.2,-0.25)$) rectangle ($(trgr) + (2,0.25)$) node[pos=.5,align=center] {\scriptsize{frame 2}};
+      \draw<7-> ($(trgr) + (0.2,.25)$) rectangle ($(trgr) + (2,.75)$) node[pos=.5,align=center] {\scriptsize{frame 3}};
+
+      \draw<7-> ($(trgr) + (1.1,1.25)$) node[align=center] {\footnotesize{Builtin call stack}};
+      \draw<7-> ($(trgr) + (0.2,-.75)$) -- ($(trgr) + (0.2,-1)$);
+      \draw<7-> ($(trgr) + (2,-.75)$) -- ($(trgr) + (2,-1)$);
+      \fill<7->[white,opacity=0.75] ($(trgr) + (-.1,1.5)$) rectangle($(trgr) + (2.5,-1.25)$);
+
+      % Higher compilations
+      \draw<8> ($(srcm) + (0.2,0.65)$) edge[-stealth,out=-90,in=45,very thick,align=center] ($(srcm) + (-0.3,-0.9)$);
     \end{tikzpicture}
@@
 
-\begin{frame}
-  \frametitle{Fully-abstract overlay semantics}
-  \begin{columns}
-\begin{column}{0.5\textwidth}
-  \begin{itemize}[<+->]
-  \item Compile from oLCM to LCM
-  \item Compilation is the identity
-  \item Builtin call stack abstraction in oLCM preserved after compilation
-  \end{itemize}
-\end{column}
-\begin{column}{0.5\textwidth}
-  \begin{tikzpicture}
-    \draw (0,0) node[draw=black, rounded corners] (olcm) {oLCM};
-    \draw node[draw=black, rounded corners,below=2.5cm of olcm] (lcm) {LCM};
-    \draw (olcm) edge[bend left, ->,thick] (lcm);
-    \draw node[left = 3cm of olcm] (ll1) {};
-    \draw node[right = 3cm of olcm] (rl1) {};
-    \draw node[below = 1.25cm of ll1] (ll) {};
-    \draw node[below = 1.25cm of rl1] (rl) {};
-    \draw[dashed] (ll) -- (rl);
+% \begin{frame}
+%   \frametitle{Fully-abstract overlay semantics}
+%   \begin{columns}
+% \begin{column}{0.5\textwidth}
+%   \begin{itemize}[<+->]
+%   \item Compile from oLCM to LCM
+%   \item Compilation is the identity
+%   \item Builtin call stack abstraction in oLCM preserved after compilation
+%   \end{itemize}
+% \end{column}
+% \begin{column}{0.5\textwidth}
+%   \begin{tikzpicture}
+%     \draw (0,0) node[draw=black, rounded corners] (olcm) {oLCM};
+%     \draw node[draw=black, rounded corners,below=2.5cm of olcm] (lcm) {LCM};
+%     \draw (olcm) edge[bend left, ->,thick] (lcm);
+%     \draw node[left = 3cm of olcm] (ll1) {};
+%     \draw node[right = 3cm of olcm] (rl1) {};
+%     \draw node[below = 1.25cm of ll1] (ll) {};
+%     \draw node[below = 1.25cm of rl1] (rl) {};
+%     \draw[dashed] (ll) -- (rl);
 
-    \draw<1> (olcm) edge[bend left, ->,thick] node[right, near start] { } (lcm);
-    \draw<2-> (olcm) edge[bend left, ->,thick] node[right, near start] {$id$} (lcm);
-  \end{tikzpicture}
-\end{column}
-\end{columns}
-\end{frame}
+%     \draw<1> (olcm) edge[bend left, ->,thick] node[right, near start] { } (lcm);
+%     \draw<2-> (olcm) edge[bend left, ->,thick] node[right, near start] {$id$} (lcm);
+%   \end{tikzpicture}
+% \end{column}
+% \end{columns}
+% \end{frame}
+
+% \begin{frame}
+%   \frametitle{Evaluating the definition}
+%   \begin{enumerate}
+%   \item {\itshape Intuitive}\\
+%     \onslide<2->{Yes, call stack corresponds exactly to our intuition.}
+% \item {\itshape Useful for reasoning}\\% we should be able to use WBCF and LSE
+%                                 % when reasoning about correctness and security
+%                                 % of programs using \stktokens{}.
+%   \onslide<3->{Yes, ...}
+% \item {\itshape Reusable in secure compiler chains}\\% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
+%   \onslide<4->{Yes, fully-abstract compilations compose vertically, so oLCM can
+%     be used as a target for other compilation phases.}
+% \item {\itshape Arguably "complete"}\\% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
+%   \onslide<5->{Yes, ...}
+% %\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
+% \end{enumerate}
+% \end{frame}
+
+% \begin{frame}
+%   \frametitle{The \stktokens{} calling convention}
+%   \begin{onslide}<2->
+%     ... is explained in the paper.
+%   \end{onslide}
+% \end{frame}
 
 \begin{frame}
-  \frametitle{Evaluating the definition}
-  \begin{enumerate}
-  \item {\itshape Intuitive}\\
-    \onslide<2->{Yes, call stack corresponds exactly to our intuition.}
-\item {\itshape Useful for reasoning}\\% we should be able to use WBCF and LSE
-                                % when reasoning about correctness and security
-                                % of programs using \stktokens{}.
-  \onslide<3->{Yes, ...}
-\item {\itshape Reusable in secure compiler chains}\\% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
-  \onslide<4->{Yes, fully-abstract compilations compose vertically, so oLCM can
-    be used as a target for other compilation phases.}
-\item {\itshape Arguably "complete"}\\% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
-  \onslide<5->{Yes, ...}
-%\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{The \stktokens{} calling convention}
-  \begin{onslide}<2->
-    ... is explained in the paper.
-  \end{onslide}
+  \frametitle{Paper overview}
+\begin{itemize}
+\item LCM: A formalization of a simple CHERI-like capability machine with linear capabilities
+\item \stktokens{}, a new calling convention that provably guarantees LSE and WBCF on LCM
+\item A new way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics}
+\item Proof of LSE and WBCF which includes
+  \begin{itemize}
+  \item oLCM: an overlay semantics for LCM with built-in LSE and WBCF
+  \item proving full-abstraction for the embedding of oLCM into LCM by
+  \item defining and using a cross-language, step-indexed, Kripke logical relation with recursive worlds
+  \end{itemize}
+\end{itemize}
+  % \begin{description}
+  % \item[Formalization of CHERI-like capability machine with linear capabilities]
+  % \item[\stktokens{}] a calling convention that provably guarantees LSE and well-bracketed control-flow WBCF
+  % \item[Fully-abstract overlay semantics] a novel way to prove LSE and WBCF claims
+  %   \begin{itemize}
+  %   \item Proven with cross-language logical relations
+  %   \end{itemize}
+  %   % TODO Mention the proof and logical relation, move to end?
+  % \end{description}
 \end{frame}
@@
 
+\begin{frame}
+  \frametitle{Desired properties of LSE and WBCF definition}
+  \begin{enumerate}
+\item {\itshape intuitive} \label{def-prop:intuitive}
+\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
+\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
+\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
+\label{def-prop:complete}
+\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
+\end{enumerate}
+\end{frame}
+
 \end{document}

### 2019-01-14 presentations/popl19-01-2019/illu-preamble.tex
Presentation update

--- 2017-12-18 presentations/illustrations/preamble.tex
@@
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=0.9cm] {};
     \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5] {\scriptsize #4};
   }{
@@
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=0.9cm] {};
     \draw[fill=white, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5] {\scriptsize #4};
     \endonlyenv
@@
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=0.9cm] {};
     \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5] {\scriptsize #4};
   }{
@@
     \draw[fill=white] #2 rectangle #3 node[pos=.5] {};
-    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=1cm] {};
+    \draw ($#2 + (3,1.5)$) node[devil,mirrored,minimum size=0.9cm] {};
     \draw[fill=white, fill opacity=0.4] #2 rectangle #3 node[pos=.5] {};
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5] {\scriptsize #4};
     \endonlyenv
@@
   \ifthenelse{\equal{#1}{}}{
-    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw[fill=white] #2 rectangle #3 node[pos=.5] {\scriptsize #4};
   }{
@@
     \draw[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw #2 rectangle #3 node[pos=.5] {\scriptsize #4};
   }{
@@
     \draw<#1>[fill=gray, fill opacity=0.5] #2 rectangle #3 node[pos=.5] {};
-    \draw<#1> #2 rectangle #3 node[pos=.5] {\footnotesize #4};
+    \draw<#1> #2 rectangle #3 node[pos=.5] {\scriptsize #4};
   }
@@
 \newcommand{\capbrace}[3][sp1]{
-  \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+  \draw [decorate,decoration={brace,amplitude=8pt,mirror,raise=3pt},yshift=0pt]
   #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
@@
     \draw (6,0) -- (6,15);
-    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+    \draw[fill=gray!50] (0,-.5) rectangle (6,2.5) node[pos=.5] {\scriptsize lower stack frames...};
   \endscope
-  \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+  % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+  \draw (-1.5,0) node {};
+  \draw (-1.5,15) node {};
 }

### 2019-01-14 presentations/popl19-01-2019/presentation.tex
Presentation update

@@
 \input{preamble}
-\input{../illustrations/preamble}
+\input{illu-preamble}
 
@@
 
-\AtBeginSection[]
-{
-  \begin{frame}<beamer>
-    \frametitle{Overview}
-    \tableofcontents[currentsection]
-  \end{frame}
-}
+% \AtBeginSection[]
+% {
+%   \begin{frame}<beamer>
+%     \frametitle{Overview}
+%     \tableofcontents[currentsection]
+%   \end{frame}
+% }
 
@@
 
-\begin{frame}
-  \frametitle{Overview}
-  \tableofcontents
-\end{frame}
+% \begin{frame}
+%   \frametitle{Overview}
+%   \tableofcontents
+% \end{frame}
 
@@
 
-\input{../illustrations/traditional}
-\input{../illustrations/capabilities}
+\begin{frame}
+  \centering
+  \frametitle{Traditional stack pointers}
+  \begin{tikzpicture}[scale=.45, every node={scale=.45}]
+    % recurrent parts
+    \begin{scope}
+      \clip (-.1,-.1) rectangle (6.1,15.1);
+      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
+      \draw[draw=gray!80] (0,0) -- (0,15);
+      \draw[draw=gray!80] (6,0) -- (6,15);
+      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\scriptsize lower stack frames...};
+    \end{scope}
+    \draw (-1.5,0) node {};
+    \draw (-1.5,15) node {};
+    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
+
+    % traditional:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\scriptsize caller stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
+
+    % traditional call:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\scriptsize caller stack frame};
+    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\scriptsize callee stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
+
+    % traditional return:
+    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\scriptsize caller stack frame};
+    \draw<.> (8,7) node[draw=black] (sp0) {};
+    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);
+
+    % traditional call, attack:
+    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\scriptsize caller stack frame};
+    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\scriptsize callee stack frame};
+    \draw<.-> (8,7) node[draw=black] (sp0) {};
+    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };
+
+    % read other stack frames
+    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
+    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
+    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
+
+    % break well-bracketedness
+    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
+    \draw<.-> (8,7) node[draw=black] (sp0) {};
+    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
+    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
+      skip caller frame};
+  \end{tikzpicture}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Naive stack and return capabilities}
+  \centering
+  \begin{tikzpicture}[scale=.45, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % first step: capabilities
+    \begin{onlyenv}<+-+(4)>
+      \draw[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize current stack frame};
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(3)>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack capability };
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+      \draw (7.5,4) node[right,fill=gray!50] (rp1l) {return capability data};
+    \end{onlyenv}
+
+    % prepare call: store old return capability
+    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\tiny old return cap code} (6,6.5);
+    \begin{onlyenv}<+-+(2)>
+      \draw[fill=gray!50] (0,5.5) rectangle node {\tiny old return cap data} (6,6);
+      % \draw [decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
+      % (0,0) -- (0,16) node (oldframe1) [black,midway,xshift=-0.5cm] {};
+      % \draw (0,5.75) edge[->,out=180,in=225] (oldframe1);
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
+      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
+      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return capability data};
+    \end{onlyenv}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { new stack capability };
+    \end{onlyenv}
+
+    % call using capabilities
+    \begin{onlyenv}<+->
+      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+      \draw[fill=white] (0,6.5) rectangle (6,9.5) node[pos=.5] {\footnotesize callee stack frame};
+    \end{onlyenv}
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Attack on naive stack and return capabilities }
+  \centering
+  \begin{tikzpicture}[scale=.45, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % attacker...
+    \begin{onlyenv}<+-+(2)>
+      \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
+      \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % stores stack pointer in heap
+    \begin{onlyenv}<+->
+      \begin{scope}
+        \fill[gray!20] (18,6) rectangle (22,13);
+        \draw (18,6) -- (18,13);
+        \draw (22,6) -- (22,13);
+      \end{scope}
+      \draw (20,14) node {heap memory};
+      \draw[fill=gray!50] (18,8) rectangle node[color=red] {\tiny copy of old sp} (22,8.5);
+    \end{onlyenv}
+    \draw<.> (18,8.25) edge[->,red,thick,bend left] (sp1);
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (13.5,6.5) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+      \draw (18,8.25) edge[->,red,thick,bend left] (spold);
+    \end{onlyenv}
+
+    % attacker returns
+    \begin{onlyenv}<+>
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
+    \end{onlyenv}
+
+    % attacker gets called again + uses old stack pointer.
+    \begin{onlyenv}<+>
+      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
+      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
+      \draw (18,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
+      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
+    \end{onlyenv}
+  \end{tikzpicture}
+\end{frame}
+
 \begin{frame}
   \frametitle{\stktokens{} prevent the attack}
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+  \centering
+  \begin{tikzpicture}[scale=.45, every node={scale=.5}]
     % recurrent parts
     \stdstackstart
-    \draw (16.5,5.5) node {register file};
+    \draw (17.5,5.5) node {register file};
     \begin{scope}
-      \clip (15.9,1.1) rectangle (17.1,5.1);
+      \clip (16.9,1.1) rectangle (18.1,5.1);
       \foreach \x in {1,2,...,4}
       {
-        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=teal] {};
+        \draw[fill=white] (17,\x) rectangle (18,\x+1) node[pos=.5,color=teal] {};
       };
     \end{scope}
-    \node (r1) at (16.75,4.5) {};
-    \node (r2) at (16.75,3.6) {};
+    \node (r1) at (17.75,4.5) {};
+    \node (r2) at (17.75,3.6) {};
 
@@
       \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,thick,bend left] (sp1);
+      \draw (r1) edge[->,red,thick,bend left,dashed] (sp1);
     \end{onlyenv}
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,thick,bend left] (sp2);
+      \draw (r2) edge[->,red,thick,bend left, dashed] (sp2);
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,thick,bend left] (usp);
+      \draw (r1) edge[->,red,thick,bend left,dashed] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left] (usp);
+      \draw (r1) edge[->,red,bend left,dashed] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
     \end{onlyenv}
-    \draw<.-> (r2) edge[->,red,bend left] (sp2);
+    \draw<.-> (r2) edge[->,red,bend left,dashed] (sp2);
 
@@
     \begin{onlyenv}<.>
-      \draw (14,10.25) edge[->,red,thick,bend right] (usp);
+      \draw (14,10.25) edge[->,red,thick,bend right,dashed] (usp);
       \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
@@
     \begin{onlyenv}<+->
-      \draw (14,10.25) edge[->,red,bend right] (usp);
+      \draw (14,10.25) edge[->,red,bend right,dashed] (usp);
       \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
@@
 
+\begin{frame}
+  \frametitle{FAQ}
+  \begin{itemize}
+  \item \textit{Do you support tail calls?}
+    \begin{itemize}
+    \item Yes.
+    \end{itemize}
+  \item \textit{Do you support higher-order functions?}
+    \begin{itemize}
+    \item Yes.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \end{document}

### 2019-01-14 presentations/popl19-01-2019/presentation.tex
Correct date in POPL presentation
POPL presentation update

@@
 
+\tikzset{
+  lock/.pic = {
+    \draw[fill=black] (0,0) rectangle (.2,.2);
+    \draw[even odd rule,fill=black] (.1,.2) circle[radius=.08]
+    (.1,.2) circle[radius=.06];
+    \draw[fill=white](.086,.054) rectangle (.114,.14);
+    \draw[fill=white,draw=none](.1,.13) circle[radius=.02];
+  }
+}
+
+% \draw <coord> pic {lock};
+
 % \AtBeginSection[]
@@
 
-\date{POPL, January 19, 2019}
+\date{POPL, January 16, 2019}
 
@@
       \draw node[rectangle,fill=white,draw] at (1.93,3.7) (ir1) {...};
-      \draw node[rectangle,fill=white,rounded corners,draw] at (3.15,3.7) (ir2) {ir};
+      \draw node[rectangle,fill=white,rounded corners,draw] at (3.15,3.7) (ir2) {IR};
       \draw (0.8, 3.7) edge[-stealth,bend left,very thick] node[above] {} (ir1);
@@
   \frametitle{\stktokens{} prevent the attack}
+  % Emphazise that it is dynamically enforced.
+  % Other cases, see paper.
   \centering
@@
 
-    \begin{onlyenv}<.->
+    \begin{onlyenv}<+->
       \begin{scope}
@@
     % Callee stores stack pointer
-    \begin{onlyenv}<+->
+    \begin{onlyenv}<.->
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
 \begin{frame}[fragile]
-  \frametitle{Fully-abstract overlay Semantics}
+  \frametitle{Fully-abstract overlay semantics}
 \onslide<8->{\hdashrule[0.1ex]{12cm}{1pt}{5pt}}

### 2019-01-15 presentations/popl19-01-2019/presentation.tex
POPL19 presentation update

@@
   lock/.pic = {
-    \draw[fill=black] (0,0) rectangle (.2,.2);
-    \draw[even odd rule,fill=black] (.1,.2) circle[radius=.08]
+    \draw[even odd rule,fill=gray] (.1,.2) circle[radius=.08]
     (.1,.2) circle[radius=.06];
-    \draw[fill=white](.086,.054) rectangle (.114,.14);
-    \draw[fill=white,draw=none](.1,.13) circle[radius=.02];
+    \draw[fill=gray] (0,0) rectangle (.2,.2);
+
+    \draw (.086,.054) rectangle (.114,.14);
+    \draw (.1,.13) circle[radius=.03];
+
+    \draw[fill=white,draw=none](.096,.064) rectangle (.104,.13);
+    \draw[fill=white,draw=none] (.1,.13) circle[radius=.02];
   }
@@
     \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
-    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
+    \draw<.> (sp0) edge[->,color=myred,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);
 
     % break well-bracketedness
-    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
+    \draw<+>[fill=myred,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
     \draw<.-> (8,7) node[draw=black] (sp0) {};
-    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
-    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
+    \draw<.>[myred] (sp0) edge[->,out=235,in=0] (6,1.75);
+    \draw<.> node[color=myred,align=left,below=of sp0l] {return but\\
       skip caller frame};
@@
       \draw (7.5,4) node[right,fill=gray!50] (rp1l) {return capability data};
+      \draw (7.4,4.3) pic {lock};
     \end{onlyenv}
@@
       \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return capability data};
+      \draw ($(rp1l.west) + (-.1,.3)$) pic {lock};
     \end{onlyenv}
@@
       (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
-      \draw node[right=0cm of sp1] { new stack capability };
+      \draw node[right=0cm of sp1] (nsp) { new stack capability };
     \end{onlyenv}
@@
       \draw (20,14) node {heap memory};
-      \draw[fill=gray!50] (18,8) rectangle node[color=red] {\tiny copy of old sp} (22,8.5);
+      \draw[fill=gray!50] (18,8) rectangle node[color=myred] {\tiny copy of old sp} (22,8.5);
     \end{onlyenv}
-    \draw<.> (18,8.25) edge[->,red,thick,bend left] (sp1);
+    \draw<.> (18,8.25) edge[->,myred,thick,bend left] (sp1);
     \begin{onlyenv}<+->
@@
       (13.5,6.5) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
-      \draw (18,8.25) edge[->,red,thick,bend left] (spold);
+      \draw (18,8.25) edge[->,myred,thick,bend left] (spold);
     \end{onlyenv}
@@
       \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
-      \draw (18,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+      \draw (18,8.25) edge[->,color=myred,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
       \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
       {
-        \draw[fill=white] (17,\x) rectangle (18,\x+1) node[pos=.5,color=teal] {};
+        \draw[fill=white] (17,\x) rectangle (18,\x+1) node[pos=.5,color=mygreen] {};
       };
@@
       \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,red,thick,bend left,dashed] (sp1);
+      \draw (r1) edge[->,myred,thick,bend left,dashed] (sp1);
     \end{onlyenv}
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,red,thick,bend left, dashed] (sp2);
+      \draw (r2) edge[->,myred,thick,bend left, dashed] (sp2);
       \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,thick,bend left,dashed] (usp);
+      \draw (r1) edge[->,myred,thick,bend left,dashed] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,red,bend left,dashed] (usp);
+      \draw (r1) edge[->,myred,bend left,dashed] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
-      \draw node[right=0cm of sp2] { return data pointer };
+      \draw node[right=0cm of sp2] (rdp) { return data pointer };
+      \draw ($(rdp.west) + (-.1,.3)$) pic {lock};
       \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
@@
     \end{onlyenv}
-    \draw<.-> (r2) edge[->,red,bend left,dashed] (sp2);
+    \draw<.-> (r2) edge[->,myred,bend left,dashed] (sp2);
 
@@
     \end{onlyenv}
-    \draw<.>[fill=white] (14,10) rectangle node[color=red] {} (18,10.5);
+    \draw<.>[fill=white] (14,10) rectangle node[color=myred] {} (18,10.5);
 
@@
     \begin{onlyenv}<.>
-      \draw (14,10.25) edge[->,red,thick,bend right,dashed] (usp);
-      \draw[fill=white] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
+      \draw (14,10.25) edge[->,myred,thick,bend right,dashed] (usp);
+      \draw[fill=white] (14,10) rectangle node[color=myred] {\tiny old sp} (18,10.5);
       \draw node[right=0cm of usp] { stack pointer };
@@
       \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
-      \draw node[right=0cm of sp2] { return data pointer };
+      \draw node[right=0cm of sp2] (rdp2) { return data pointer };
+      \draw ($(rdp2.west) + (-.1,.3)$) pic {lock};
     \end{onlyenv}
@@
     \begin{onlyenv}<+->
-      \draw (14,10.25) edge[->,red,bend right,dashed] (usp);
-      \draw[fill=gray!50] (14,10) rectangle node[color=red] {\scriptsize old sp} (18,10.5);
+      \draw (14,10.25) edge[->,myred,bend right,dashed] (usp);
+      \draw[fill=gray!50] (14,10) rectangle node[color=myred] {\tiny old sp} (18,10.5);
       \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
@@
 
-    \draw<+>[teal,very thick] (6.4,4.4) -- (7.4,5.4);
-    \draw<.>[teal,very thick] (7.4,4.4) -- (6.4,5.4);
-    \draw<.> node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
+    \draw<+>[mygreen,very thick] (6.4,4.4) -- (7.4,5.4);
+    \draw<.>[mygreen,very thick] (7.4,4.4) -- (6.4,5.4);
+    \draw<.> node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,4) { Splice fails,\\ stack pointer is linear!};  
   \end{tikzpicture}  
@@
 \item \stktokens{}, a new calling convention that provably guarantees LSE and WBCF on LCM
-\item A new way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics}
+\item A better way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics}
 \item Proof of LSE and WBCF which includes
@@
 
-\end{document}
+\end{document}p

### 2019-01-15 presentations/popl19-01-2019/presentation.tex
Change return data pointer name to return token

@@
 
-      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
-      \draw node[right=0cm of sp2] (rdp) { return data pointer };
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return token} };
+      \draw node[right=0cm of sp2] (rdp) { return token };
       \draw ($(rdp.west) + (-.1,.3)$) pic {lock};
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return data pointer} };
-      \draw node[right=0cm of sp2] (rdp2) { return data pointer };
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return token} };
+      \draw node[right=0cm of sp2] (rdp2) { return token };
       \draw ($(rdp2.west) + (-.1,.3)$) pic {lock};

### 2019-01-22 toplas/paper.tex
typo
Toplas paper wording.
Implement some of TOPLAS reviewer 1's comments

@@
 \begin{toplas}
-  This paper is an extension of a published conference paper\citep{Skorstengaard:esop18}.
+  This paper is an extension of a published conference paper~\citep{Skorstengaard:esop18}.
   We have made minor improvements to readability and completeness throughout the paper, but we highlight the following changes:
   \begin{itemize}
-  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provide informal intuition about how the logical relation machinery comes into play on a capability machine.
-  \item We have added a proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
+  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
+  \item We have added proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
   \item We have added figures that illustrate central parts of the calling convention.
@@
   \item We have added a section on reasoning about programs that run on a capability machine.
-    This section explains how one reason about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
+    This section explains how one reasons about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
     It also introduces a number of lemmas that prove recurring bits once and for all.
@@
     &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
-     && \Caps  &::= & \multicolumn{6}{l}{\left\{
+     && \Caps  &\isdef & \multicolumn{6}{l}{\left\{
                          ((\perm,\gl),\start,\addrend,\addr)\mid
@@
 words, our worlds are defined recursively. Recursive worlds are common
-in Kripke models and the following lemma uses the method of
+in Kripke models and the following theorem uses the method of
 \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
@@
 transition system.  The different monotonicity requirements in the two
-interpretation functions reflects how permanent regions rely only on permanent
+interpretation functions reflect how permanent regions rely only on permanent
 protocols whereas temporary regions can rely on both temporary and permanent
@@
 \forall (n, ms) \in A. \forall m
-\leq n. (m, ms) \in A\}$. 
+\leq n\ldotp (m, ms) \in A\}$. 
 \
@@
 \begin{toplas}
-  In the future world relations, we must remember all region names in order to
-  keep them extensional which is why we need to use masks. It does, however, not
-  matter what kind of region is used to mask as it is the region name we must
-  retain. This is why we allow revoked regions to be replaced by other
-  regions in the two future world relations. This may seem unnecessary as the
-  future world relation allows new regions to be added, but in private future
-  worlds, for technical reasons, we need to be able to reinstate old regions
-  with the same region names as they had in a past world. Further, we do not
-  want public future worlds to be able to prevent reinstation in private future
-  worlds. Private future worlds cannot revoke permanent regions, so we cannot
-  allow public future worlds to us permanent regions as a mask for a revoked
-  region.
+  In order to be able to solve the recursive domain equation, we want the future world relations to be extensional.
+  This means a future world cannot forget about region names in a past world.
+  This is the reason we use masks rather than deleting the regions altogether.
+  It does not, however, matter what kind of region is used as a mask; it is the region name we must retain.
+  This is why we allow revoked regions to be replaced by other regions in the two future world relations.
+  This may seem unnecessary as the future world relation allows new regions to be added, but in private future worlds, for technical reasons, we need to be able to reinstate old regions with the same region names as they had in a past world.
+  Further, we do not want public future worlds to be able to prevent reinstation in private future worlds.
+  Private future worlds cannot revoke permanent regions, so we cannot allow public future worlds to us permanent regions as a mask for a revoked region.
 \end{toplas}
@@
 As described previously, the specification of malloc ensures that malloc has no
-capabilities pointing \emph{out} of malloc. It does, however, not say anything about
+capabilities pointing \emph{out} of malloc. It does not, however, say anything about
 capabilities that point \emph{in} to malloc. We obviously cannot

### 2019-01-24 toplas/paper.tex
Implement reviewer 1 comments

@@
 % \usepackage[numbers]{natbib}
+\usepackage{xcolor}
 \usepackage{url}
@@
   keepspaces=true,
-  mathescape
+  mathescape,
+  escapeinside=||,
+  numbers=left,
+  numberblanklines=false,
+%  frame=single,
+%  framexleftmargin=12pt,
+  numbersep=5pt,
+  morecomment=[l]{//},
+  commentstyle=\color{blue},
+  morekeywords={push, pop, move, lea, restrict, subseg, geta, getb, mclear,
+    rclear, jmp, load, store, fetch, scall, assert, halt, malloc, call, crtcls,
+    reqglob, prepstk, plus, minus, jnz, gete}
 }
@@
 \newenvironment{toplas}{}{}
-%    {\color{OliveGreen}}{}
+    {\color{OliveGreen}}{}
 \newenvironment{toplassug}
@@
     H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  H^\pwl_A \; s \; \hat{W} \defeq{} & \left\{\npair{\hs} \middle|
+  H^\pwl_A (s)(\hat{W}) \defeq{} & \left\{\npair{\hs} \middle|
     \begin{aligned}
-      &\dom(\hs) = A \land \\
+      &\dom(\hs) = A \text{ and } \\
       &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
@@
     H^\nwl_A :{} & \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  H^\nwl_A \; s \;\hat{W} \defeq{} & \left\{\npair{\hs} \middle|
+  H^\nwl_A (s)(\hat{W}) \defeq{} & \left\{\npair{\hs} \middle|
     \begin{aligned}
-      &\dom(\hs) = A \land \\
+      &\dom(\hs) = A \text{ and } \\
       &\forall \addr \in A \ldotp \\
-      &\quad \nonlocal{\ms(\addr)} \land\\ 
+      &\quad \nonlocal{\ms(\addr)} \text{ and }\\ 
       &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
@@
 The \emph{register-file relation} $\stdrr$ defines safe register-files as those
-that contain safe words (i.e.\ words in $\stdvr$) in all registers but $\pcreg$.
+that contain safe words (i.e.\ words in the $\stdvr$-relation defined below) in all registers but $\pcreg$.
 The \emph{expression relation} $\stder$ defines that a word is safe to use as a
@@
 value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{[\start,\addrend]}$
-means that $W(r)$ must accept only safe values like
-$\iota^\pwl_{[\start,\addrend]}$, but can be even more restrictive if desired. The
+means that $W(r)$ must accept only memory segments containing safe values like
+$\iota^\pwl_{[\start,\addrend]}$, but can be even more restrictive if needed
+(the $\nsubsim[n]$-relation is defined in Appendix~\ref{app:logical-relation}). The
 read condition also takes into account the locality of the capability because,
@@
 $\iota^\pwl_{[\start,\addrend]}$ can be found
-in \figurename~\ref{fig:std-reg}; it makes use of the isomorphism from
+in \figurename~\ref{fig:std-reg}; it makes use of the isomorphism $\xi$ from
 Theorem~\ref{thm:world-existence}.
@@
 two accepted memory segments. An interesting property of the write condition is
-that they prohibit global write-local capabilities which, as discussed in
+that it prohibits global write-local capabilities which, as discussed in
 \sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
@@
             \text{If }
-            \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
-            \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
+            \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land{} \\
+            \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land{} \\
             \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
@@
             \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
-            \qquad \quad \Phi \step[j] \Phi' \land \\
-            \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
-            \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
-            \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
-            \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
-            \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
+            \qquad \quad \Phi \step[j] \Phi' \land{} \\
+            \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land{}\\
+            \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land{} \\
+            \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land{} \\
+            \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land{} \\
+            \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land{} \\
             \qquad \quad \size - 1 = e'-b'
@@
 pushed to the stack before the call are popped to the private registers. The
-implementation of \texttt{scall} is presented in
-\figurename~\ref{fig:scall-impl}, and the restoration code is presented in
-\figurename~\ref{fig:scall-rest-code}. The implementation of \texttt{scall} uses
+implementation of \texttt{scall} and the restoration code used by
+\texttt{scall} are presented in \figurename~\ref{fig:scall-impl}. The implementation of \texttt{scall} uses
 some of the macros we present next.
-
-\begin{figure}[tb]
-    \centering
-\begin{lstlisting}
- $i_1$ = move r_t1 pc
- $i_2$ = lea r_t1 $\var{ofc}$
- $i_3$ = load r_stk r_t1
- $i_4$ = pop pc
-\end{lstlisting}
-    \caption{The restoration code used in \texttt{scall}. The variable
-      $\var{ofc}$ is 5 which is the offset to the address where the old stack
-      pointer is stored on the stack.}
-    \label{fig:scall-rest-code}
-\end{figure}
 
@@
     \centering
+    \begin{minipage}[t]{.6\linewidth}
 \begin{lstlisting}
-// push private registers to the stack
+// Push the private registers to the stack.
   push $r_{\var{priv},1}$
@@
   push $r_{\var{priv},n}$
-// push restoration code to the stack
+// Push the restoration code to the stack.
   push $\encode(i_1)$
@@
   push $\encode(i_4)$
-// push old pc to the stack
+// Push the old pc to the stack.
   move r_t1 pc
@@
   push r_t1
-// push stack pointer to the stack
+// Push the stack pointer to the stack.
   push r_stk
-// set up protected return pointer
+// Set up the protected return pointer.
   move r_0 r_stk
-  lea r_0 $\var{ofc_\var{rec}}$ // -5 is the offset to the first instruction of
-the recovery code
+  lea r_0 $\var{ofc_\var{rec}}$
   restrict r_0 $\encodePermPair((\local,\entry))$
-// restrict stack capability to unused part
+// Restrict the stack capability to the unused part of the stack.
   geta r_t1 r_stk
@@
   subseg r_stk r_t1 r_t2
-// clear unused part of the stack
+// Clear the unused part of the stack.
   mclear r_stk
-// clear non-argument registers
+// Clear non-argument registers.
   rclear $R$
   jmp r
-after:
-// pop the restore code
+return:
+// Pop the restore code.
   pop r_t1
@@
   pop r_t1
-// pop the private state into approriate registers
+// Pop the private state into approriate registers.
   pop $r_{\var{priv},1}$
@@
 \end{lstlisting}      
+    \end{minipage}
+    \begin{minipage}[t]{.29\linewidth}
+\begin{lstlisting}[title=\footnotesize{The restoration code used in \texttt{scall}.}, captionpos=b,xleftmargin=2em,frame=single,framexleftmargin=1.5em]
+ $i_1$ = move r_t1 pc
+ $i_2$ = lea r_t1 $\var{ofc}_\var{stk}$
+ $i_3$ = load r_stk r_t1
+ $i_4$ = pop pc
+\end{lstlisting}
+    \end{minipage}%
     \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
       r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
-      restoration code is given in \figurename~\ref{fig:scall-rest-code}. The
-      variable $\var{ofc}$ is the offset to the label after, and
+      restoration code is presented in the top right corner. The
+      variable $\var{ofc}_\var{ret}$ is the offset to the label \texttt{return}, and
       $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
       the activation record. The set $R = \RegName \setminus
-      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.}
+      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.
+    The variable $\var{ofc}_\var{stk} =5$ which is the offset to the address
+    where the old stack pointer is stored on the stack.}
     \label{fig:scall-impl}
@@
 closure's code and a new environment is allocated (using malloc) and the
-contents of $\overline{r_i}$ is stored in the environment. In the code referred
+contents of $\overline{r_i}$ are stored in the environment. In the code referred
 to by $r$, an implicit load from the environment happens when an instruction
@@
 carry over to proofs about assembly programs, so many of the proofs in the next
-section about example programs are a bit cumbersome. It is especially annoying,
+section about example programs are a bit cumbersome. It is especially annoying
 when the same line of reasoning is applied in multiple places. To mitigate this,
@@
   \label{lem:scall-works}
-  If
+  For all $n \in \nats$, $\ms, \ms_\stk, \ms_\var{unused}, \ms_f \in \MemSegments$, $W
+  \in \Worlds$, $\reg \in \Regs$, $r,\overline{r_{\mathit{arg}}},
+  \overline{r_{\mathit{priv}}} \in \RegName$, and $c_\var{next} \in \Caps$,
+  if
   \begin{enumerate}
@@
     \overline{r_{\mathit{priv}}}) \text{ followed by }
-    c_{\mathit{next}}$\footnote{Defined in the appendix.}
+    c_{\mathit{next}}$
   \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
-  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} If
+  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} For all  $\ms_\var{rec},
+    \ms_\var{unused}', \ms'' \in \MemSegments$, $W' \in \Worlds$, $c_\var{ret},
+    c_\stk \in \Caps$, and $\reg' \in \Regs$, if
     \begin{itemize}
@@
     then we have that $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
-  \item \label{item:scall:hyp-cont} \emph{Hyp-Cont} If
+  \item \label{item:scall:hyp-cont} \emph{Hyp-Cont} For all $n' \in \nats$, $W''
+    \in \Worlds$, $\ms'', \ms_\var{unused}''$, and $\reg' \in \Regs$, if
     \begin{itemize}
@@
       \end{equation*}
-    \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$ for some $\ms_{\mathit{unused}}''$
+    \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$
     \end{itemize}
@@
 \end{lemma}
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}{0.43\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+      \scope
+      \clip (-.1,-.1) rectangle (4.6,11.1);
+      \fill[fill=white] (0,0) rectangle (4.5,11);
+      \fill[color=gray!20] (0,0) rectangle (4.5,4);
+      \fill[color=gray!20] (0,9.5) rectangle (4.5,11);
+      \draw[color=gray!50] (0,5.5) rectangle (4.5,6);
+      \draw[color=gray!50] (0,7) rectangle (4.5,7.5);
+      \draw (0,0) -- (0,11);
+      \draw (4.5,0) -- (4.5,11);
+      \draw (0,4) -- (4.5,4);
+      \draw (0,9.5) -- (4.5,9.5);
+      \endscope
+      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+      \draw (-1.5,0) node {};
+      \draw (-1.5,15) node {};
+
+      \node at (8.5,7.5) { $\reg$ };
+      \begin{scope}
+        \clip (7.9,4.5) rectangle (9.1,7.1);
+        \foreach \x in {4,5,...,6}
+        {
+          \draw (8,\x) rectangle (9,\x + 1);
+        };
+      \end{scope}
+      \node[anchor=west] at (9.125,6.5) {\footnotesize\texttt{pc}};
+
+      \node at (2.25,7.25) { \footnotesize{$i_0$} };
+      \node at (2.25,6.68) { \footnotesize{$\vdots$} };
+      \node at (2.25,5.75) { \footnotesize{$i_n$} };
+      \node[rotate=0] at (2.25,8.5) {\huge{$\ms$}};
+
+      \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.59,mirror},yshift=0pt]
+      (4.6,4) -- (4.6,9.5) node (cstk) [black,midway,xshift=0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.23,mirror},yshift=0pt]
+      (5,4) -- (5,9.5) node (cstk) [black,midway,xshift=0.8cm] {};
+
+      \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
+            {\fill[black] circle(1.5pt);}}}}
+      \draw[capa] (8.5,6.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\tiny{$(g,l)$}} (5,7.25);
+
+      \node[anchor=west] at (7.75,2) {$c_\var{next}$};
+      \draw[capa] (7.75,2) edge[-latex,postaction={decorate},out=180,in=0] node[right] {\tiny{$(g,l)$}}  (5.4,5.26);
+    \end{tikzpicture}
+    \caption{Visualization of ``$(\reg,\ms)$ is looking at $[i_0\dots i_n]$
+      followed by $c_{\var{next}}$''.}
+  \label{fig:def-intuition-looking-at}
+  \end{subfigure}
+  \begin{subfigure}{.12\linewidth}
+$\phantom{te}$
+  \end{subfigure}
+  \begin{subfigure}{0.43\linewidth}
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+      \scope
+      \clip (-.1,-.1) rectangle (4.6,11.1);
+      \fill[fill=white] (0,0) rectangle (4.5,11);
+      \fill[color=gray!20] (0,9.5) rectangle (4.5,11);
+      \fill[color=gray!20] (0,0) rectangle (4.5,2);
+      \draw (0,0) -- (0,11);
+      \draw (4.5,0) -- (4.5,11);
+
+      \draw (0,9.5) -- (4.5,9.5);
+      \draw (0,4) -- (4.5,4);
+      \draw (0,2) -- (4.5,2);
+      \endscope
+      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+      \draw (-1.5,0) node {};
+      \draw (-1.5,15) node {};
+
+      \node at (8.5,7.5) { $\reg$ };
+      \begin{scope}
+        \clip (7.9,4.5) rectangle (9.1,7.1);
+        \foreach \x in {4,5,...,6}
+        {
+          \draw (8,\x) rectangle (9,\x + 1);
+        };
+      \end{scope}
+      \node[anchor=west] at (9.125,5.5) {\footnotesize\texttt{r\_stk}};
+
+      \node[rotate=0] at (2.25,7) {\Large{$\ms_{\stk,\var{unused}}$}};
+      \node[rotate=0] at (2.25,3) {\Large{$\ms_{\stk}$}};
+
+      \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.23,mirror},yshift=0pt]
+      (4.6,2) -- (4.6,9.5) node (cstk) [black,midway,xshift=0.8cm] {};
+
+      \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
+            {\fill[black] circle(1.5pt);}}}}
+      \draw[capa] (8.5,5.5) edge[-latex,postaction={decorate},out=200,in=10,densely dashed] node[above,sloped] {\tiny $\rwlx$} (5,3.75);
+
+    \end{tikzpicture}
+    \caption{Visualization of ``$\reg$ points to stack with $\ms_\stk$ used and
+      $\ms_{\stk,\var{unused}}$ unused''.}
+  \label{fig:def-intuition-points-to-stack}
+  \end{subfigure}
+
+  \caption{
+  }
+  \label{fig:def-intuition}
+\end{figure}
+\begin{toplas}
+The \texttt{scall} lemma is stated in terms of a number of auxiliary definitions
+found in the appendix. The region $\iota^{sta}(\gl,\ms)$ is a static region with
+locality $\gl$. It is static in the sense that it only accepts the region $\ms$
+(\appendixname~\ref{app:awkward-example}).
+The function $\revokeTemp{} : \Worlds \fun \Worlds$ yields the input world but
+with all the temporary regions replaced with revoked regions (\appendixname~\ref{app:logical-relation}). The definition of \textit{$(\reg,\ms)$ is looking at $[i_0\dots i_n]$
+      followed by $c_{\var{next}}$} is visualized in
+    \figurename~\ref{fig:def-intuition-looking-at};
+    it requires the capability in $\pcreg$ of
+    $\reg$ to point to the first address of the instructions $[i_0\dots i_n]$ in
+    $\ms$ and $c_{\var{next}}$ to point to the address immediately after the
+    instructions. The definition of \textit{$\reg$ points to stack with $\ms_\stk$ used and
+      $\ms_{\stk,\var{unused}}$ unused} is visualized in
+    \figurename~\ref{fig:def-intuition-points-to-stack};
+    it requires $\ms_\stk$ and $\ms_{\stk,\var{unused}}$ to be adjacent and
+    continuous memory segments with the local $\rwlx$-capability in
+    \texttt{r\_stk} of $\reg$ governing the two memory segments and pointing to
+    the top most address of $\ms_\stk$ (\appendixname~\ref{app:programs-definitions}).
+\end{toplas}
+
 Roughly, the \texttt{scall} lemma states that an invocation of scall is safe if
@@
   \begin{minipage}[t]{.37\linewidth}
-\begin{lstlisting}
+\begin{lstlisting}[firstnumber=last]
      ${\text{\tiny\itshape(continued from previous column)}}$
@@
   \begin{minipage}[t]{.25\linewidth}
-  \begin{lstlisting}
+  \begin{lstlisting}[firstnumber=last]
 f3: push 1
@@
   \inferrule{(s,\phi_\pub,\phi) = (s',\phi_\pub',\phi') \\
-    \forall \hat{W} \ldotp H \; s \; \hat{W} \nsubeq H' \; s' \; \hat{W} }
+    \forall \hat{W} \ldotp H \; (s) \; (\hat{W}) \nsubeq H' \; (s') \; (\hat{W}) }
   { (v,s,\phi_\pub,\phi,H) \nsubsim (v',s',\phi_\pub',\phi',H')}
@@
   \activeReg{W} & \defeq \dom(\erase{W}{\perma,\temp})
+\end{align*}
+%
+\textbf{Revoke temporary regions in a world}
+\begin{align*}
+  \revokeTemp{} & : \Worlds \fun \Worlds \\
+  \revokeTemp{W} & \defeq \lambda r \ldotp 
+                   \begin{cases}
+                     \revoked            & \text{if }W(r) = (\temp,s,\phi_\pub,\phi,H) \\
+                     W(r)                & \text{otherwise}
+                   \end{cases}
 \end{align*}
@@
 \subsection{Reasoning about programs definitions}
+\label{app:programs-definitions}
 \begin{definition}
@@
 \begin{definition}
-  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c_\malloc}'' 
+  We say that ``\linksto{(\reg,\ms)}{\var{key}}{j}{c}'' 
   iff
@@
 \subsection{Awkward example}
+\label{app:awkward-example}
 \textbf{The region for variable $x$} The region $\iota_x$, is the region omitted from the proof sketch for the awkward example.
@@
     \begin{aligned}
-      &\ms' = \ms \land\\
+      &\ms' = \ms \land{}\\
       &\forall \addr \in \dom(\ms) \ldotp\\
-      & \quad \nonlocal{\ms(\addr)} \land\\
+      & \quad \nonlocal{\ms(\addr)} \land{}\\
       & \quad \npair[n-1]{\ms(\addr)} \in \stdvr(\xi(\hat{W}))

### 2019-01-24 presentations/popl19-01-2019/presentation.tex

@@
 {
+  int t = 5;
   printf("Hello, World!");
+
   return 0;
@@
   \frametitle{Fully-abstract overlay semantics}
-\onslide<8->{\hdashrule[0.1ex]{12cm}{1pt}{5pt}}
+\onslide<7->{\hdashrule[0.1ex]{12cm}{1pt}{5pt}}
 \begin{tabular}{p{3.2cm} p{3.2cm}}
@@
 
-      % Call stack
-      \begin{scope}
-        \clip<7-> ($(trgr) + (0.1,-1.2)$) rectangle ($(trgr) + (2.1,-.75)$);
-        \draw<7-> ($(trgr) + (0.2,-1.5)$) rectangle ($(trgr) + (2,-.75)$) node[pos=.5,align=center] {\scriptsize{...}};
-      \end{scope}
-      \draw<7-> ($(trgr) + (0.2,-0.75)$) rectangle ($(trgr) + (2,-0.25)$) node[pos=.5,align=center] {\scriptsize{frame 1}};
-      \draw<7-> ($(trgr) + (0.2,-0.25)$) rectangle ($(trgr) + (2,0.25)$) node[pos=.5,align=center] {\scriptsize{frame 2}};
-      \draw<7-> ($(trgr) + (0.2,.25)$) rectangle ($(trgr) + (2,.75)$) node[pos=.5,align=center] {\scriptsize{frame 3}};
+      % % Call stack
+      % \begin{scope}
+      %   \clip<7-> ($(trgr) + (0.1,-1.2)$) rectangle ($(trgr) + (2.1,-.75)$);
+      %   \draw<7-> ($(trgr) + (0.2,-1.5)$) rectangle ($(trgr) + (2,-.75)$) node[pos=.5,align=center] {\scriptsize{...}};
+      % \end{scope}
+      % \draw<7-> ($(trgr) + (0.2,-0.75)$) rectangle ($(trgr) + (2,-0.25)$) node[pos=.5,align=center] {\scriptsize{frame 1}};
+      % \draw<7-> ($(trgr) + (0.2,-0.25)$) rectangle ($(trgr) + (2,0.25)$) node[pos=.5,align=center] {\scriptsize{frame 2}};
+      % \draw<7-> ($(trgr) + (0.2,.25)$) rectangle ($(trgr) + (2,.75)$) node[pos=.5,align=center] {\scriptsize{frame 3}};
 
-      \draw<7-> ($(trgr) + (1.1,1.25)$) node[align=center] {\footnotesize{Builtin call stack}};
-      \draw<7-> ($(trgr) + (0.2,-.75)$) -- ($(trgr) + (0.2,-1)$);
-      \draw<7-> ($(trgr) + (2,-.75)$) -- ($(trgr) + (2,-1)$);
-      \fill<7->[white,opacity=0.75] ($(trgr) + (-.1,1.5)$) rectangle($(trgr) + (2.5,-1.25)$);
+      % \draw<7-> ($(trgr) + (1.1,1.25)$) node[align=center] {\footnotesize{Builtin call stack}};
+      % \draw<7-> ($(trgr) + (0.2,-.75)$) -- ($(trgr) + (0.2,-1)$);
+      % \draw<7-> ($(trgr) + (2,-.75)$) -- ($(trgr) + (2,-1)$);
+      % \fill<7->[white,opacity=0.75] ($(trgr) + (-.1,1.5)$) rectangle($(trgr) + (2.5,-1.25)$);
 
       % Higher compilations
-      \draw<8> ($(srcm) + (0.2,0.65)$) edge[-stealth,out=-90,in=45,very thick,align=center] ($(srcm) + (-0.3,-0.9)$);
+      \draw<7> ($(srcm) + (0.2,0.65)$) edge[-stealth,out=-90,in=45,very thick,align=center] ($(srcm) + (-0.3,-0.9)$);
     \end{tikzpicture}

### 2019-01-24 toplas/paper.tex
Addressed more of TOPLAS reviewer 2 comments

@@
   For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
-  Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{c.f.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
+  Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{cf.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
   This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.
-  In $W_2$, we replace the stack region of $W_1$ with a temporary static region that governs our stack frame, that is the private registers we push to the stack and the stack recovery code, as well as a standard region $\iota^{pwl}$-region for the part of the stack, we will provide to the callback.
-  The static region ensures that our local stack frame remains unchanged during the execution of the callback, and the standard region makes sure that the callback only puts safe values on the stack.
+  In $W_2$, two new regions govern the stack.
+  One of the new two regions governs the caller stack frame, i.e.\ the part of the stack that contains the contents of private registers and stack recovery code.
+  The region that takes care of this is a temporary static region which ensures that our local stack frame remains unchanged during the execution of the callback.
+  The other new region we add takes care of the unused part of the stack which we are going to let the callback use for its execution.
+  This is taken care of by a standard region $\iota^{pwl}$-region which allows the callback to store any safe value on the stack.
   The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
@@
   This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still present in $W_3$.
-  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 0), so for the part of the execution up until the \texttt{scall} we once again use the anti-reduction lemma.
+  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 1), so for the part of the execution up until the \texttt{scall} we once again use the anti-reduction lemma.
 
@@
 like the well-bracketedness that Example~\texttt{g1} relies on.
-In fact, our examples show that our logical relation imply a stronger form of
+In fact, our examples show that our logical relation implies a stronger form of
 control-flow correctness than such formulations, although this is not made very
@@
 Consider the awkward example proof, for example, we would have to ensure that the world $W_6$, depicted in Figure~\ref{fig:worlds-in-pf-g1} would not just be a private future world of $W_5$, but a public one (in addition to being a public future world of $W_1$).
-This would allow us to do without the stack clearing, but it would entail some changes to the islands we use, and an extra argument that the old adversary's stack and return pointer remain valid after clearing the trusted code's stack frame and relaxing the invariant that used to ensure it could not be modified.
+This would allow us to do without the stack clearing, but it would entail some changes to the regions we use, and an extra argument that the old adversary's stack and return pointer remain valid after clearing the trusted code's stack frame and relaxing the invariant that used to ensure it could not be modified.
 While this change is a clear improvement, we do not actually believe that it fundamentally changes the efficiency characteristics of the approach: the cost is halved, but remains asymptotically the same.
@@
 However, our requirement for stack clearing on boundary crossings is also general.
-Revoking authority that was granted temporarily using local capabilities, requires clearing all memory for which the invokee had write-local authority (or at least erasing all local capabilities from that memory).
+Revoking authority that was granted temporarily using local capabilities requires clearing all memory for which the invokee had write-local authority (or at least erasing all local capabilities from that memory).
 Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenarios where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
@@
 
-Note, by the way, that CheriBSD's use of local capabilities in ccall does not actually involve a form of revocation.
+Note, by the way, that CheriBSD's use of local capabilities in CCall does not actually involve a form of revocation.
 CheriBSD's model involves a trusted stack manager that gives every compartment access to its own private stack using a local, write-local capability \citep{Watson2015Cheri}.
@@
 Recent work suggests that such a feature could perhaps be added to processors like CHERI, using a special hardware cache that tracks whether or not a memory location contains zero \citep{Joannou2017EfficientTaggedMemory}.
+
 % http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf
@@
   \begin{cases}
-    \dom(\erase{W}{\perm,\temp}) & \text{if } \gl = \local \\
-    \dom(\erase{W}{\perm}) & \text{if } \gl = \glob
+    \dom(\erase{W}{\perma,\temp}) & \text{if } \gl = \local \\
+    \dom(\erase{W}{\perma}) & \text{if } \gl = \glob
   \end{cases}

### 2019-01-25 toplas/paper.tex
Address reviewer 3 concerns
Address the remaining of TOPLAS reviewer comments, first pass

@@
 \definecolor{Blue}{RGB}{92,26,143}
-\newenvironment{toplas}{}{}
+\newenvironment{toplas}%{}{}
     {\color{OliveGreen}}{}
@@
 
-\newcommand{\itoplas}[1]{#1}
-%    {{\color{OliveGreen} #1}}
+\newcommand{\itoplas}[1]%{#1}
+    {{\color{OliveGreen} #1}}
 \newcommand{\itoplassug}[1]{}
@@
 
-A similar problem is the enforcement of control-flow correctness on capability
-machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
-the standard contiguous C stack is split into a central, trusted stack, managed
-by trusted call and return instructions, and disjoint, private, per-compartment
-stacks. To prevent illegal use of stack references, the approach relies on
-\emph{local capabilities}, a type of capabilities offered by CHERI to
-\emph{temporarily} relinquish authority, namely for the duration of a function
-invocation whereafter the capability can be revoked. However, details are scarce
-(how does it work precisely? what features are supported?) and a lot remains to
-be investigated (e.g., combining disjoint stacks with cross-domain function
-pointers seems like it will scale poorly to large numbers of components?).
-Finally, there is no argument that the approach is watertight and it is not even
-clear what security property is targeted exactly.
+A similar problem is the enforcement of control-flow correctness on capability machines.
+An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}: the standard contiguous C stack is split into a central, trusted stack and disjoint, private, per-compartment stacks.
+The trusted stack is managed by trusted call and return instructions.
+To prevent illegal use of stack references, the approach relies on \emph{local capabilities}, a type of capabilities offered by CHERI to \emph{temporarily} relinquish authority, namely for the duration of a function invocation whereafter the capability can be revoked.
+However, details are scarce (how does it work precisely?
+what features are supported?)
+and a lot remains to be investigated (e.g., combining disjoint stacks with cross-domain function pointers seems like it will scale poorly to large numbers of components).
+Finally, there is no argument that the approach is watertight and it is not even clear what security property is targeted exactly.
 
@@
 
-\begin{figure}[hb]
+We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
+We assume an infinite set of addresses $\Addrs$ and define machine words as either integers or capabilities of the form $((\perm,\gl),\var{base},\var{end},\addr)$.
+Such a capability represents the authority to execute permissions $\perm$ on the memory range $[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag $\gl$ indicating whether the capability is global or local.
+There is no notion of pointers other than capabilities, so we will use the terms interchangeably.
+The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$) permissions.
+Additionally, there are three special permissions: read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter ($\entry$), which we will explain below.
+\itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}.
+  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory.
+  We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
+
+\begin{wrapfigure}{R}{.4\textwidth}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (7) {$\rwlx$};
+    \node[main node] (8) [below left of=7] {$\readwritel$};
+    \node[main node] (1) [below right of=7] {$\rwx$};
+    \node[main node] (2) [below right of=1] {$\exec$};
+    \node[main node] (3) [below right of=2] {$\entry$};
+    \node[main node] (4) [below left of=1] {$\readwrite$};
+    \node[main node] (5) [below right of=4] {$\readonly$};
+    \node[main node] (6) [below right of=5] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (7) edge (8)
+    (7) edge (1)
+    (8) edge (4)
+    (1) edge (2)
+    (2) edge (3)
+    (2) edge (5)
+    (3) edge (6)
+    (1) edge (4)
+    (4) edge (5)
+    (5) edge (6);
+
+    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
+    \node[main node] (2l) [below of=1l] {$\local$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (1l) edge (2l);
+  \end{tikzpicture}
+  \caption{Permission and locality hierarchy.}
+  \label{fig:perm-hier}
+\end{wrapfigure}
+
+We assume a finite set of register names $\RegName$. We define register files
+$\reg$ and memories $\ms$ as functions mapping register names resp. addresses to
+words. The state of the entire machine is represented as a configuration that is
+either a running state $\Phi \in \ExecConfs$ containing a memory and a register file,
+or a failed or halted state, where the latter keeps hold of the final state of
+memory.
+% We also define memory segments $\ms$:
+% parts of memory represented as partial functions mapping addresses to words.
+%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
+
+The machine's instruction set is rather basic. Instructions $i$ include
+relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
+($\move{}{}$, copies words between registers) instructions. Also familiar are
+load and store instructions for reading from and writing to memory ($\load{}{}$
+and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and
+$\minus{}{}{}$, operating only on numbers). There are three instructions for
+modifying capabilities: $\lea{}{}$ (modifies the current address),
+$\restricttwo{}{}$ (modifies the permission and local/global tag) and
+$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
+instructions take care that the resulting capability always carries less
+authority than the original (e.g. $\restricttwo{}{}$ will only weaken a permission).
+Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a
+number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and
+$\geta{}{}$ provide access to a capability's permissions, local/global tag, base,
+end and current address, respectively.
+
+\begin{figure}[b]
 \[\arraycolsep=1.4pt
@@
                  \load{\lv}{\hv} \mid 
-                 \store{\hv}{\rv} \mid \plus{\lv}{\rv}{\rv} \mid 
+                 \store{\hv}{\rv} \mid
+                 \lt{\lv}{\rv}{\rv} \mid
+                 \plus{\lv}{\rv}{\rv} \mid 
                  \minus{\lv}{\rv}{\rv} \mid \\
-          &   &  \lt{\lv}{\rv}{\rv} \mid
-                \lea{\lv}{\rv} \mid 
+          &   &  \lea{\lv}{\rv} \mid 
                  \restricttwo{\lv}{\rv} \mid
@@
                  \isptr{\lv}{\rv} \mid 
-                 \getl{\lv}{\lv} \mid \\
-          &   &  \getp{\lv}{\lv} \mid 
-                 \getb{\lv}{\lv} \mid
-                 \gete{\lv}{\lv} \mid 
+                 \getl{\lv}{\lv} \mid
+                 \getp{\lv}{\lv} \mid
+                 \getb{\lv}{\lv} \mid\\
+          &   &  \gete{\lv}{\lv} \mid 
                  \geta{\lv}{\lv} \mid 
@@
     \hline
-    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
-                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
-    \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
-    \hline
     $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$ and $(\perm',g') = \decodePermPair{\memreg(r_2)}$
@@
     \hline
+    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
+    \hline
     \itoplas{$\lea{\refreg{r_1}}{r_2}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{c}}}$} &\itoplas{$\memreg(r_1) = \stdcap$ and $n = \memreg(r_2)$ and $n \in \ints $ and $\perm \neq \entry$ and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$} \\
@@
                                                 $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
+    \hline
+    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
+                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
     \hline
@@
 \end{figure}
-
-We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
-We assume an infinite set of addresses $\Addrs$ and define machine words as
-either integers or capabilities of the form
-$((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability represents the
-authority to execute permissions $\perm$ on the memory range
-$[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag
-$\gl$ indicating whether the capability is global or local. There is no notion
-of pointers other than capabilities, so we will use the terms interchangeably.
-The available permissions are null permission
-($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
-($\exec$) and read/write/execute ($\rwx$) permissions. Additionally, there are
-three special permissions: read/write-local ($\readwritel$), read/write-local/execute
-($\rwlx$) and enter ($\entry$), which we will explain below.
-\itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}. We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
-
-\begin{wrapfigure}{R}{.4\textwidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (7) {$\rwlx$};
-    \node[main node] (8) [below left of=7] {$\readwritel$};
-    \node[main node] (1) [below right of=7] {$\rwx$};
-    \node[main node] (2) [below right of=1] {$\exec$};
-    \node[main node] (3) [below right of=2] {$\entry$};
-    \node[main node] (4) [below left of=1] {$\readwrite$};
-    \node[main node] (5) [below right of=4] {$\readonly$};
-    \node[main node] (6) [below right of=5] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (7) edge (8)
-    (7) edge (1)
-    (8) edge (4)
-    (1) edge (2)
-    (2) edge (3)
-    (2) edge (5)
-    (3) edge (6)
-    (1) edge (4)
-    (4) edge (5)
-    (5) edge (6);
-
-    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
-    \node[main node] (2l) [below of=1l] {$\local$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (1l) edge (2l);
-  \end{tikzpicture}
-  \caption{Permission and locality hierarchy.}
-  \label{fig:perm-hier}
-\end{wrapfigure}
-
-We assume a finite set of register names $\RegName$. We define register files
-$\reg$ and memories $\ms$ as functions mapping register names resp. addresses to
-words. The state of the entire machine is represented as a configuration that is
-either a running state $\Phi \in \ExecConfs$ containing a memory and a register file,
-or a failed or halted state, where the latter keeps hold of the final state of
-memory.
-% We also define memory segments $\ms$:
-% parts of memory represented as partial functions mapping addresses to words.
-%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
-
-The machine's instruction set is rather basic. Instructions $i$ include
-relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
-($\move{}{}$, copies words between registers) instructions. Also familiar are
-load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and
-$\minus{}{}{}$, operating only on numbers). There are three instructions for
-modifying capabilities: $\lea{}{}$ (modifies the current address),
-$\restricttwo{}{}$ (modifies the permission and local/global tag) and
-$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
-instructions take care that the resulting capability always carries less
-authority than the original (e.g. $\restricttwo{}{}$ will only weaken a permission).
-Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a
-number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and
-$\geta{}{}$ provide access to a capability's permissions, local/global tag, base,
-end and current address, respectively.
 
@@
 into a register, but only if the capability has appropriate authority (i.e.\
-read permission and an appropriate range). $\restricttwo{}{}$ updates a
+read permission and an appropriate range). The $\restricttwo{}{}$ instruction updates a
 capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original. 
@@
 % Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
-It also never turns local
-capabilities into global ones. $\geta{}{}$ queries the current address of a
-capability and stores it in a register.
+It also never turns local capabilities into global ones.
+The $\subseg{}{}{}$ instruction reduces the range of authority of a capability.
+In order to represent the unbounded address space, we use $-42$ to represent infinity.
 
@@
 jumped to, the instructions can start executing and have access to the local
-data through the updated version of $c$ that is then in $\pcreg$. 
+data through the updated version of $c$, now in the $\pcreg$-register. 
 \lau{As we talked about what is described here, is not how we make closures in
@@
 here. }
+
+The instruction $\lea{}{}$ manipulates the current address of non enter-capabilities.
+It is fine for a capability to have a current address outside its range of authority as long as it is not used with an instruction that requires a specific capability.
+The instruction $\geta{}{}$ queries the current address of a capability and stores it in a register.
 
@@
 \emph{temporarily} given to an adversary, for the duration of an invocation: if
-we take care to clear the capability from the register file after control is
+we make sure to clear the capability from the register file after control is
 passed back to us, they will not have been able to store the capability.
@@
 memory as follows: a contiguous array of memory is reserved for the stack, for
-which we receive a stack pointer in a special register $r_\stk$. We stress that
+which we receive a stack pointer in the register $r_\stk$. We stress that
 the stack is not built-in, but merely an abstraction we put on this piece of the
@@
 %   is so important that we do not want to leave it out.}
-By making the return pointer local, we prevent such attacks: the adversary can
-only store local capabilities through write-local capabilities, which means (because
-of our assumptions above): on the stack. Since the stack pointer itself is
-also local, it can also only be stored on the stack. Because we clear the part
-of the stack that the adversary has had access to before we pass control back,
-there is no way for them to recover either of these local capabilities.
+By making the return pointer local, we prevent such attacks.
+The adversary can only store local capabilities with a write-local capability, and the only piece of memory governed by a write-local capability is the stack.
+Since the stack pointer itself is also local, it can also only be stored on the stack.
+There is no way for the adversary to recover either of these local capabilities because we clear the part of the stack that they had access to before we pass control back to them.
 
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine the adversary stores their stack pointer (\ref{fig:illu-adv-store-sp}), invokes trusted code
-that uses part of the stack to store private data (\ref{fig:illu-adv-call}) and then invokes the adversary
+pointers. Imagine the adversary stores their stack pointer (\figurename~\ref{fig:illu-adv-store-sp}), invokes trusted code
+that uses part of the stack to store private data (\figurename~\ref{fig:illu-adv-call}) and then invokes the adversary
 again with a stack pointer restricted to exclude the part containing the private
-data (\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
+data (\figurename~\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
 could access the private data stored there by the trusted code and break
@@
 
-Clearly, neither of these capabilities should be accessible by the adversary. In
+Clearly, neither the old program counter or the old stack pointer should be accessible by the adversary. In
 other words, the return pointer provided to the adversary must be a capability
@@
 \paragraph{Local capabilities}
-A final thing we provide informal intuition for is how our results take into account local
-capabilities and their special treatment by the hardware. Normally, when we
-invoke an untrusted piece of code and provide it with certain global
-capabilities, it may have stored those capabilities in memory and we will only
-be able to reinvoke the code if we can guarantee that those values are still
-valid. Formally, worlds represent the invariants that global capabilities'
-safety relies on and the reinvocation is only safe in future worlds, where the
-invariants are respected.
-
-However, if we provide the adversary with local capabilities in that first
-invocation, then the situation is a bit different. The adversary has no way to
-store these local capabilities, so if we make sure that there are also no old
-local capabilities in the register file for the second invocation (including the
-capability being invoked), then the adversary cannot use them any more.%
-\footnote{We are ignoring write-local capabilities in this discussion. If the
-  adversary does have access to write-local capabilities in the first and second
-  invocation, then the memory they address must entirely be cleared before the
-  second invocation in order for the reinvocation to remain safe.} Therefore, we
-can allow the second invocation to happen in \emph{private} future worlds ($W'
-\futurestr W$), in which safe global capabilities remain safe, but local ones
-don't. This private future world relation is more liberal than the standard,
-public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain
-safe). Concretely, worlds may contain \emph{temporary} regions, representing
-invariants that only local capabilities may rely on for their safety, and which
-may be revoked (disabled) in private future worlds.
-
-Interestingly, this idea is a variant of a notion of public/private future
-worlds that has been previously used in the literature (see
-\sectionname~\ref{sec:discussion}). However, temporary regions are new in our setting
-and there is an interesting interplay with the recursiveness of the worlds: for
-a temporary region, the predicate $H \in \Worlds \rightarrow
-\powerset{\MemSegments}$ (which defines the safe memory segments in the
-current world) is only required to be monotone w.r.t.\ public
-future worlds (i.e.\ safe memory segments remain safe in public future worlds),
-but for permanent regions, it must be monotone w.r.t.\ private future worlds.
-As a consequence, the memory for a permanent region may not contain local
-capabilities (as their safety would be broken in private future worlds), which
-in turn implies that only local capabilities may have write-local permission (a
-general sanity requirement when using local capabilities\footnote{In fact, local
-  capabilities become useless as soon as the adversary has access to a single
-  global, write-local capability.})\footnote{For explanation purposes, this
-  discussion ignores certain ways to allow for local capabilities in a permanent
-  region, for example, by not requiring that they are valid or requiring that
-  they are local versions of valid global capabilities.}.
+The last thing we provide informal intuition about is how our results handle local capabilities and their special treatment by the hardware.
+Normally, when we invoke an untrusted piece of code and provide it with certain global capabilities, it may have stored those capabilities in memory and we will only be able to reinvoke the code if we can guarantee that those values are still valid.
+Formally, worlds represent the invariants that global capabilities' safety relies on and the reinvocation is only safe in future worlds, where the invariants are respected.
+
+However, if we provide the adversary with local capabilities in that first invocation, then the situation is a bit different.
+The adversary has no way to store these local capabilities, so if we make sure that there are also no old local capabilities in the register file for the second invocation (including the capability being invoked), then the adversary cannot use them any more.%
+\footnote{We are ignoring write-local capabilities in this discussion.
+  If the adversary does have access to write-local capabilities in the first and second invocation, then the memory they address must entirely be cleared before the second invocation in order for the reinvocation to remain safe.}
+Therefore, we can allow the second invocation to happen in \emph{private} future worlds ($W' \futurestr W$), in which safe global capabilities remain safe, but local ones don't.
+This private future world relation is more liberal than the standard, public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain safe).
+Concretely, worlds may contain \emph{temporary} regions, representing invariants that only local capabilities may rely on for their safety, and which may be revoked (disabled) in private future worlds.
+
+Interestingly, this idea is a variant of a notion of public/private future worlds that has been previously used in the literature (see \sectionname~\ref{sec:discussion}).
+However, temporary regions are new in our setting and there is an interesting interplay with the recursiveness of the worlds: for a temporary region, the predicate $H \in \Worlds \rightarrow \powerset{\MemSegments}$ (which defines the safe memory segments in the current world) is only required to be monotone w.r.t.\ public future worlds (i.e.\ safe memory segments remain safe in public future worlds), but for permanent regions, it must be monotone w.r.t.\ private future worlds.
+As a consequence, the memory for a permanent region may not contain local capabilities (as their safety would be broken in private future worlds), which in turn implies that only local capabilities may have write-local permission (a general sanity requirement when using local capabilities\footnote{In fact, local capabilities become useless as soon as the adversary has access to a single global, write-local capability.})\footnote{For explanation purposes, this discussion ignores certain ways to allow for local capabilities in a permanent region, for example, by not requiring that they are valid or requiring that they are local versions of valid global capabilities.}.
 \end{toplas}
@@
             {\fill[black] circle(1.5pt);}}}}
-      \draw[capa] (8.5,6.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\tiny{$(g,l)$}} (5,7.25);
+      \draw[capa] (8.5,6.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\tiny{$(p,\gl)$}} (5,7.25);
 
       \node[anchor=west] at (7.75,2) {$c_\var{next}$};
-      \draw[capa] (7.75,2) edge[-latex,postaction={decorate},out=180,in=0] node[right] {\tiny{$(g,l)$}}  (5.4,5.26);
+      \draw[capa] (7.75,2) edge[-latex,postaction={decorate},out=180,in=0] node[right] {\tiny{$(p,\gl)$}}  (5.4,5.26);
     \end{tikzpicture}
@@
   \begin{cases}
-    \dom(\erase{W}{\perma,\temp}) & \text{if } \gl = \local \\
+    \dom(\erase{W}{\perma ,\temp}) & \text{if } \gl = \local \\
     \dom(\erase{W}{\perma}) & \text{if } \gl = \glob

### 2019-01-31 toplas/paper.tex
Add paragraph about why we need LR

@@
 convention enforces control-flow correctness, we do not prove a general theorem
-that shows this, because it is not clear what such a theorem should look like.
+that shows this because it is not clear what such a theorem should look like.
 Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties,
@@
 
+\begin{toplas}
+  \emph{Why use a logical relation and not a simpler proof technique?}
+  The concept of a capability exists on different levels of the computer stack, so capabilities have been studied and safety properties proven about them in other contexts than assembly languages.
+  Traditionally, logical relations have not been used for safety property proofs, so why do we need one here?
+  To answer this question let's compare this work to \citet{sewel_integrity_2011} where they prove \textit{integrity} and \textit{authority confinement} for the seL4 Microkernel.
+  % Describe what they prove
+  The security properties they show are relative to a security policy that specifies the upperbound of capabilities a subject in the system can posses.
+  The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject a can have a write capability for memory B.}
+  However, they cannot express what \emph{subject a} will be able to store in \emph{memory B}.
+  The worlds in our model can put restrictions on how a capability can be used by placing a protocol on part of the memory.
+  This is exactly what allows us to prove the awkward example.
+We place a protocol on the memory where variable \texttt{x} is stored that says that it sometimes can be either $0$ or $1$, and at other times it has to be $1$.
+\end{toplas}
 \itoplassug{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}

### 2019-01-31 toplas/paper.tex
Added comment on our notion of capability safety considers meaning of programs

@@
   We place a protocol on the memory where variable \texttt{x} is stored that says that it sometimes can be either $0$ or $1$, and at other times it has to be $1$.
+  Capability safety is not just about the permission a capability carries, it is also about how the capability is used.
+  Say for instance, a program has a write capability that will only be used to write even numbers.
+  If this is the only write capability for that part of memory, then we should be able to rely on that part of memory only containing even numbers.
+  In order to rely on this in our proofs, our notion of capability safety must be able to take the meaning of the program into account and express the memory invariant.
 \end{toplas}

### 2019-02-01 toplas/paper.tex
Remove reference to missing figure
Update seL4 paragraph

@@
 We stress that the macros correspond to series of instructions as seen in
-\figurename~\ref{fig:scall-impl} and \ref{fig:scall-rest-code}, and that they
+\figurename~\ref{fig:scall-impl}, and that they
 are introduced for intelligibility. The examples of
@@
   The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject a can have a write capability for memory B.}
-  However, they cannot express what \emph{subject a} will be able to store in \emph{memory B}.
-  The worlds in our model can put restrictions on how a capability can be used by placing a protocol on part of the memory.
-  This is exactly what allows us to prove the awkward example.
+  However, their policy are, in a sense, binary: either subject a can not write to memory B, or it can write anything.
+  Our approach is much more fine-grained, allowing us to define, for example, a policy that subject a can write to memory B, but only if the value being written is an even number.
+  Because our invariants are indexed by worlds themselves, we can even define a policy that capabilities written to memory B must respect certain other invariants themselves.
+  It is exactly the expressive invariants that enable us to prove the awkward example.
   We place a protocol on the memory where variable \texttt{x} is stored that says that it sometimes can be either $0$ or $1$, and at other times it has to be $1$.
+
   Capability safety is not just about the permission a capability carries, it is also about how the capability is used.

### 2019-02-02 toplas/paper.tex
Change State to WState

@@
 \newcommand{\Rels}{\plaindom{Rels}}
-\newcommand{\States}{\plaindom{State}}
+\newcommand{\States}{\plaindom{WState}}
 \newcommand{\RegionNames}{\plaindom{RegionName}}
@@
 state transition system where $\Rels$ contains pairs of relations corresponding
-to the public and private transitions, and $\States$ is an unspecified set that
-we assume to contain at least the states we use in this paper. The last part of
+to the public and private transitions, and $\States$ is a set of world states that
+we assume to at least contain the states we use in this paper. The last part of
 the temporary and permanent regions is a state interpretation function that

### 2019-02-03 toplas/paper.tex
Include call in appendix and mention as part of heap based CC

@@
 \subsection{Encapsulation of Local State}
-\texttt{f1} and \texttt{f2} in
-\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability
-machine's encapsulation of local state. They are very similar: both
-store some local state, call an untrusted piece of code ($\adv$), and
-then test whether the local state is unchanged. They differ in the way
-they do this. Program \texttt{f1} uses our stack-based
-calling convention (captured by \texttt{scall}) to call
-the adversary, so it can use the available stack to store its local
-state.  On the other hand, \texttt{f2} uses malloc to
-allocate memory for its local state and uses a calling convention based on heap
-allocated activation records (described in the technical appendix) to invoke
-the adversarial code.
+\texttt{f1} and \texttt{f2} in The programs in \figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's encapsulation of local state.
+They are very similar: both store some local state, call an untrusted piece of code ($\adv$), and then test whether the local state is unchanged.
+They differ in the way they do this.
+Program \texttt{f1} uses our stack-based calling convention (captured by \texttt{scall}) to call the adversary, so it can use the available stack to store its local state.
+On the other hand, \texttt{f2} uses malloc to allocate memory for its local state and uses \texttt{call} a calling convention based on heap allocated activation records (described in Appendix~\ref{app:macros}) to invoke the adversarial code.
 
@@
 \subsection{Macros}
+\label{app:macros}
 Implementation of the macros used in \texttt{scall}. Implementations of the
@@
 appendix~\citep{technical_appendix}.\\
-\label{app:macros}
 \texttt{push $r$}
-\begin{lstlisting}
+\begin{lstlisting}[xleftmargin=0.4cm]
   lea r_stk 1
@@
 \texttt{pop $r$}
-\begin{lstlisting}
+\begin{lstlisting}[xleftmargin=0.4cm]
   load r r_stk
@@
 \texttt{rclear $r_1,\dots, r_n$}
-\begin{lstlisting}
+\begin{lstlisting}[xleftmargin=0.4cm]
   move $r_1$ 0
@@
 \texttt{mclear $r$}
-\begin{lstlisting}
+\begin{lstlisting}[xleftmargin=0.4cm]
   move r_t $r$
@@
 \end{lstlisting}
-Where $\var{ofc}_\var{end}$ and $\var{ofc}_\var{iter}$ are the offsets to the
-label \texttt{end} % 9?
-and \texttt{iter}, % 2?
+Where $\var{ofc}_\var{end}$ and $\var{ofc}_\var{iter}$ are the offsets to the label \texttt{end} % 9
+and \texttt{iter}, % 2
 respectively.
+\newline\newline
+\texttt{call} $r(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$
+\newline
+The \texttt{call} macro constitutes a calling convention based on heap allocated activation records.
+This alternative to \texttt{scall} is included to illustrate that the logical relation can be used to reason about other calling conventions.
+In the following, $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
+\begin{itemize}
+\item Set up activation record
+\item Create local enter capability for activation (protected return pointer)
+\item Clear unused registers
+\item Jump
+\item Upon return: Run activation code
+  \begin{itemize}
+  \item Restore private registers
+  \item Jump to return capability
+  \end{itemize}
+\end{itemize}
+\begin{lstlisting}[xleftmargin=0.4cm]
+  malloc r_t $size$ 
+// store private state in activation record
+  store r_t r_priv,1
+  lea r_t 1
+  store r_t r_priv,2
+  lea r_t 1
+  ...
+  lea r_t 1
+  store r_t r_priv,n
+  lea r_t 1
+// store old pc
+  move r_t1 pc
+  lea r_t1 $\var{off}_\var{end}$ 
+  store r_t r_t1
+  lea r_t1 1
+// store activation record
+  store r_t encode(i_1)
+  lea r_t1 1
+  ...
+  lea r_t1 1  
+  store r_t encode(i_m)
+  lea r_t1 $k$ 
+  restrict r_t1 encodePermPair(($\local$,e))
+  move r_0 r_t1
+// Clear unused registers
+  rclear R // R = RegisterName - {r,pc,r_0,r_args}
+  jmp r
+end:
+\end{lstlisting}
+Where $\bar{r_{\var{priv}}} = r_{\var{priv},1}, \dots, r_{\var{priv},n}$, $size$ is the size of the activation record, $\var{off}_\var{end}$ is the offset to the \texttt{end} label, and $k$ is $m-1$, i.e.
+the offset to the first instruction of the activation code.\newline
+The activation record. The instructions correspond to $i_1,\dots,i_m$ in the above.
+\begin{lstlisting}[xleftmargin=0.4cm]
+  move r_t pc
+  getb r_t1 r_t
+  geta r_t2 r_t
+  minus r_t1 r_t1 r_t2
+// load private state
+  lea r_t r_t1
+  load r_priv,1 r_t
+  lea r_t 1
+  load r_priv,2 r_t
+  lea r_t 1
+  ...
+  lea r_t 1
+  load r_priv,n r_t
+  lea r_t 1
+// load old pc
+  load pc r_t
+\end{lstlisting}
 

### 2019-02-03 toplas/paper.tex
Add comment about temporal properties of CFG approach

@@
 that shows this because it is not clear what such a theorem should look like.
-Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties,
-like the well-bracketedness that Example~\texttt{g1} relies on.
+Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
+Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
 In fact, our examples show that our logical relation implies a stronger form of

### 2019-02-03 toplas/paper.tex
Add mention of scalability of proofs

@@
 \begin{toplas}
-In the future we would also like to investigate a new program logic for
+In the future, we would also like to investigate a new program logic for
 reasoning about capability safety for our capability machine model. In fact,
 we think the lemmas in Section~\ref{sec:reasoning} are suggestive of the style
-of results that could be written in such a logic. We think Iris would also be
-a natural starting point for such an endeavour, since Iris is really a
-framework, which can be instantiated to different programming languages. OCPL
-was able to leverage existing Iris specifications for the high-level language;
+of results that could be captured in such a logic. We think Iris would also be
+a natural starting point for such an endeavour since Iris is a
+framework that can be instantiated with different programming languages. OCPL
+was able to leverage existing Iris specifications for a high-level language;
 for our capability machine model, however, it would be necessary to devise new
@@
 logics for assembly programming languages, such as XCAP~\citep{xcap}.
+Building a logic around the semantic model presented here would remove some of the tedious repetitive proof details making it more realistic to prove properties about larger more realistic programs.
 \end{toplas}

### 2019-02-04 toplas/paper.tex
Update the introduction

@@
 
-Compromising software security is often based on attacks that break programming
-language properties relied upon by software authors, such as control-flow
-correctness, local-state encapsulation, etc. Commodity processors offer little
-support for defending against such attacks: they offer security primitives with
-only coarse-grained memory protection and limited compartmentalization
-scalability. As a result, defenses against attacks on control-flow correctness
-and local-state encapsulation are either limited to only certain common forms of
-attacks (leading to an attack-defense arms race) and/or rely on techniques like
-machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code
-verification~\citep{morrisett_system_1999}, virtual machines with a
-native stack~\citep{lindholm_java_2014} or
-randomization~\citep{forrest_building_1997}. The latter techniques essentially
-emulate protection techniques on existing hardware, at the cost of performance,
-system complexity and/or security.
+Compromising software security is often based on attacks that break programming language properties relied upon by software authors, such as control-flow correctness, local-state encapsulation, etc.
+Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
+As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-mitigation arms race) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
+The latter techniques essentially emulate protection techniques on existing hardware, at the cost of performance, system complexity and/or security.
 
@@
 (ideally evidenced by a mathematical proof, guaranteeing that we do not end up
-in a new attack-defense arms race). However, a lot needs to happen before we get
+in a new attack-mitigation arms race). However, a lot needs to happen before we get
 there. For example, it is far from trivial to devise a compilation scheme
@@
 \begin{toplas}
-  This paper is an extension of a published conference paper~\citep{Skorstengaard:esop18}.
-  We have made minor improvements to readability and completeness throughout the paper, but we highlight the following changes:
+  This paper is an extension of the published conference paper~\citet{Skorstengaard:esop18}.
+  We have made improvements to readability and completeness throughout the paper, but we highlight the following changes:
   \begin{itemize}
-  \item We have added an introduction to the \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
-  \item We have added proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
-  \item We have added figures that illustrate central parts of the calling convention.
-  \item We have added a section on malloc.
+  \item An introduction to the \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
+  \item Proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
+  \item Figures that illustrate central parts of the calling convention.
+  \item A section about malloc.
     Specifically, we provide the specification for the malloc used in the examples of the paper.
-  \item We have added a section on macro instructions with descriptions of all the macros and the implementation of \texttt{scall}.
-  \item We have added a section on reasoning about programs that run on a capability machine.
+  \item A section about macro instructions with descriptions of all the macros and the implementation of \texttt{scall}.
+  \item A section on reasoning about programs that run on a capability machine.
     This section explains how one reasons about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
     It also introduces a number of lemmas that prove recurring bits once and for all.
-  \item We have added many details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
-    Further, many of the details we did still not include in the paper are now in the appendix of this paper.
+  \item Details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
+    Further, many of the still omitted details have been added to the paper's appendix.
   \end{itemize}

### 2019-02-05 toplas/paper.tex
Minor textual updates

@@
 \label{sec:capab-mach-with}
-In this paper, we work with a formal capability machine with all the
-characteristics of real capability machines, as well as local capabilities much
-like CHERI's. Otherwise, it is kept as simple as possible. It is inspired by
-both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and
-CHERI~\citep{Watson2015Cheri}. To avoid uninteresting details, we assume an
-infinite address space and unbounded integers.
+In this paper, we work with a formalization of a capability machine with all the characteristics of real capability machines as well as local capabilities much like CHERI's.
+Otherwise, it is kept as simple as possible.
+It is inspired by both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and CHERI~\citep{Watson2015Cheri}.
+To avoid tedious details irrelevant to the focus of this paper, we assume an infinite address space and unbounded integers.
 
@@
 Such a capability represents the authority to execute permissions $\perm$ on the memory range $[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag $\gl$ indicating whether the capability is global or local.
-There is no notion of pointers other than capabilities, so we will use the terms interchangeably.
-The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$) permissions.
+On a capability machine, there is no notion of pointers other than capabilities, so we will use the terms interchangeably.
+The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$), and read/write/execute ($\rwx$) permissions.
 Additionally, there are three special permissions: read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter ($\entry$), which we will explain below.
-\itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}.
-  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory.
+\itoplas{The orderings of permissions and locality are displayed in \figurename~\ref{fig:perm-hier}.
+  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory which is why write-local is above write in the permission hierarchy.
   We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
@@
 
-We assume a finite set of register names $\RegName$. We define register files
-$\reg$ and memories $\ms$ as functions mapping register names resp. addresses to
-words. The state of the entire machine is represented as a configuration that is
-either a running state $\Phi \in \ExecConfs$ containing a memory and a register file,
-or a failed or halted state, where the latter keeps hold of the final state of
-memory.
+We assume a finite set of register names $\RegName$.
+We define register files $\reg$ and memories $\ms$ as functions mapping register names resp.\ addresses to words.
+The state of the entire machine is represented as a configuration that is either a running state $\Phi \in \ExecConfs$ containing a memory and a register file or a failed or halted state, where the latter is paired with the final state of memory.
+
 % We also define memory segments $\ms$:
 % parts of memory represented as partial functions mapping addresses to words.
+
 %\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
 
-The machine's instruction set is rather basic. Instructions $i$ include
-relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$) and move
-($\move{}{}$, copies words between registers) instructions. Also familiar are
-load and store instructions for reading from and writing to memory ($\load{}{}$
-and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and
-$\minus{}{}{}$, operating only on numbers). There are three instructions for
-modifying capabilities: $\lea{}{}$ (modifies the current address),
-$\restricttwo{}{}$ (modifies the permission and local/global tag) and
-$\subseg{}{}{}$ (modifies the range of a capability). Importantly, these
-instructions take care that the resulting capability always carries less
-authority than the original (e.g. $\restricttwo{}{}$ will only weaken a permission).
-Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a
-number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and
-$\geta{}{}$ provide access to a capability's permissions, local/global tag, base,
-end and current address, respectively.
+The machine's instruction set is rather basic.
+Instructions $i$ include relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$), and move ($\move{}{}$, copies words between registers) instructions.
+Also familiar are load and store instructions for reading from and writing to memory ($\load{}{}$ and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and $\minus{}{}{}$, operating only on numbers).
+There are three instructions for modifying capabilities: $\lea{}{}$ (modifies the current address), $\restricttwo{}{}$ (modifies the permission and local/global tag), and $\subseg{}{}{}$ (modifies the range of a capability).
+Importantly, these instructions take care that the resulting capability always carries less authority than the original (e.g.\ $\restricttwo{}{}$ will only weaken a permission according to the hierarchy in \figurename~\ref{fig:perm-hier}).
+Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and $\geta{}{}$ provide access to a capability's permissions, local/global tag, base, end and current address, respectively.
 
@@
 In general, should we add all the missing definitions and make this document closer to being self-contained.}
-shows the operational semantics for a
-few representative instructions. Essentially, a configuration $\Phi$ either
-decodes and executes the instruction at $\memreg(\pcreg)$ if it is executable
-and its address is in the valid range or otherwise fails. The table in the
-figure shows for instructions $i$ the result of executing them in configuration
-$\Phi$. The instructions $\fail$ and $\halt$ obviously fail and halt respectively. $\move{}{}$
-simply modifies the register file as requested and updates the $\pcreg$ to the
-next instruction using the meta-function $\stdUpdatePc{}$.
+shows the operational semantics for a few instructions representative of the machine.
+Essentially, a configuration $\Phi$ either decodes and executes the instruction pointed to by $\memreg(\pcreg)$ if it is an executable capability with its address in the valid range; otherwise it fails.
+The table in the figure shows for instructions $i$ the result of executing them in configuration $\Phi$.
+The instructions $\fail$ and $\halt$ obviously fail and halt respectively.
+$\move{}{}$ simply modifies the register file as requested and updates the $\pcreg$ to the next instruction using the meta-function $\stdUpdatePc{}$.
 
@@
 read permission and an appropriate range). The $\restricttwo{}{}$ instruction updates a
-capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original. 
+capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original according to the permission hierarchy in \figurename~\ref{fig:perm-hier}.
 %\itoplassug{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
@@
 % Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
-It also never turns local capabilities into global ones.
 The $\subseg{}{}{}$ instruction reduces the range of authority of a capability.
@@
 The $\jmp{}$ instruction updates the program counter to a requested location,
-but it is complicated by the presence of \emph{enter capabilities}, modeled
+but it is complicated by the presence of \emph{enter capabilities} 
 after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
@@
 watertight, potentially efficient, does not rely on a trusted central stack
-manager and supports higher-order interfaces to an adversary, where an adversary
-is just some unknown piece of code. In this section, we explain this
-convention's high-level approach, the security measures to be taken in a number
-of situations (motivating each separately with a summary table at the end).
+manager, and supports higher-order interfaces to an adversary, where an adversary
+is just some unknown piece of code. In this section, we explain the high-level
+approach of this calling convention. We motivate each security measures with a
+situation we want to avoid (motivating each measure separately with a summary table at the end).
 After that, we define a number of reusable macro-instructions that can be used
@@
 the machine. In fact, an adversary can apply the same safety measures to protect
-themselves against us. In the next paragraphs, we will explain the issues to be
+themselves against us. In the following paragraphs, we will explain the issues to be
 considered in all the relevant situations: when (1) starting our program, (2)
@@
 \textbf{Returning to the adversary} Perhaps the simplest situation is returning
-to the adversary after they invoked our code (\figurename~\ref{fig:illu-intro}). In this case, we have received a
+to the adversary after they invoked our code \itoplas{(\figurename~\ref{fig:illu-intro})}. In this case, we have received a
 return pointer from them, and we just need to jump to it as usual. An obvious
@@
 functions etc.), so we also need to clear that data before returning to the
-adversary (\figurename~\ref{fig:illu-priv-data} and \figurename~\ref{fig:illu-clear-priv}).
+adversary \itoplas{(\figurename~\ref{fig:illu-priv-data} and \figurename~\ref{fig:illu-clear-priv})}.
 
 However, if we are returning from a function that has itself invoked adversary
-code (\figurename~\ref{fig:illu-call-adv}), then clearing the used part of the stack is not enough. The \emph{unused}
+code \itoplas{(\figurename~\ref{fig:illu-call-adv})}, then clearing the used part of the stack is not enough. The \emph{unused}
 part of the stack may also contain data and capabilities, left there by the
@@
 back. In this case, the adversary could use the unused part of the stack to
-store local pointers and load them from there after they get control back (\figurename~\ref{fig:illu-adv-data}). To
+store local pointers and load them from there after they get control back
+\itoplas{(\figurename~\ref{fig:illu-adv-data})}. To
 prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
 the stack that the adversary has had access to, not just the parts that we have
-used ourselves (\figurename~\ref{fig:illu-clear-all}). Since we may be talking about a large part of memory, this
+used ourselves \itoplas{(\figurename~\ref{fig:illu-clear-all})}. Since we may be talking about a large part of memory, this
 requirement is the most problematic aspect of our calling convention for
@@
 
-\begin{figure}[htb]
+\begin{figure}[tb]
   \centering
-  \begin{subfigure}{0.32\linewidth}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{The stack just before we return to the adversary. Our stack frame
+      is the topmost, and we do not have access to the adversary stack frame.}
   \label{fig:illu-intro}
   \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
+\begin{subfigure}{0.043\linewidth}
+  % padding
+  \phantom{a}
+\end{subfigure}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-  \caption{}
+    \caption{The stack just after we returned to the adversary but did not take
+      care to clear our local stack frame.}
   \label{fig:illu-priv-data}
   \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
+\begin{subfigure}{0.043\linewidth}
+  % padding
+  \phantom{a}
+\end{subfigure}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-  \caption{}
+    \caption{The stack just after we returned to the adversary and did take
+      care to clear our local stack frame.}
   \label{fig:illu-clear-priv}
   \end{subfigure}
-
-  \begin{subfigure}{0.32\linewidth}
+  \begin{subfigure}[t]{0.33\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-  \caption{}
+  \caption{The stack after an adversary has called us and we have called the adversary.}
   \label{fig:illu-call-adv}
 \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
+\begin{subfigure}[t]{0.043\linewidth}
+  % padding
+  \phantom{a}
+\end{subfigure}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-  \caption{}
+  \caption{The stack after the adversary has returned to us, and we have
+    returned from the first adversary call. At this point, the adversary has
+    access to the local state from the nested adversary call.}
   \label{fig:illu-adv-data}
 \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
+\begin{subfigure}[t]{0.043\linewidth}
+  % padding
+  \phantom{a}
+\end{subfigure}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-  \caption{}
+  \caption{The stack after the adversary has returned to us, and we have
+    returned from the first adversary call after we took care to clear the
+    unused part of the stack.}
   \label{fig:illu-clear-all}
   \end{subfigure}
-  \caption{Illustration of the situations related to stack clearing when returning to an adversary.
-    The greyed out areas are parts of the stack that are not accessible at the time.}
-  \label{fig:ret-adv}
+  \caption{Depictions of the stack in relation to stack clearing.
+    The greyed out areas are parts of the stack that are not accessible at the
+    time.
+    The sp capability is the stack pointer.}
+  \label{fig:ret-adv2}
 \end{figure}
@@
 
-In one of the registers, we also provide a return pointer, which must be a local
+In one of the registers, we also provide a return pointer which must be a local
 capability. If it were global, the adversary would be able to store away the
@@
 circumstances where this should not be possible. For example, they could store
-the return pointer, legally jump to it a first time, wait to be invoked again
+the return pointer, legally jump to it a first time, wait to be invoked again,
 and then jump to the old return pointer a second time, instead of the new return
 pointer received for the second invocation. Similarly, they could store the
-return pointer, invoke a function in our code, wait for us to invoke them again
+return pointer, invoke a function in our code, wait for us to invoke them again,
 and then jump to the old return pointer rather than the new one, received for the
@@
 dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine the adversary stores their stack pointer (\figurename~\ref{fig:illu-adv-store-sp}), invokes trusted code
-that uses part of the stack to store private data (\figurename~\ref{fig:illu-adv-call}) and then invokes the adversary
+pointers. Imagine that the adversary stores their stack pointer
+(\figurename~\ref{fig:illu-adv-store-sp}), invokes trusted code
+that uses part of the stack to store private data
+(\figurename~\ref{fig:illu-adv-call}) and then invokes the adversary
 again with a stack pointer restricted to exclude the part containing the private
@@
   \centering
-  \begin{subfigure}{0.32\linewidth}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{The stack after the adversary has its stored the stack pointer in
+      far up on the stack.}
   \label{fig:illu-adv-store-sp}
 \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
+\begin{subfigure}{0.043\linewidth}
+  % padding
+  \phantom{a}
+\end{subfigure}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
       \begin{scope}
-        \clip (5.6,5) rectangle (9,11.5);
-        \capbrace{(5.6,5)}{(5.6,12)}{sp'}
+        \clip (5.1,5) rectangle (9,11.5);
+        \capbrace{(5.1,5)}{(5.1,12)}{sp'}
       \end{scope}
@@
     \end{tikzpicture}
-  \caption{}
+  \caption{The stack after the adversary have called some trusted code.}
   \label{fig:illu-adv-call}
   \end{subfigure}
-  \begin{subfigure}{0.333\linewidth}
+\begin{subfigure}{0.043\linewidth}
+  % padding
+  \phantom{a}
+\end{subfigure}
+  \begin{subfigure}[t]{0.3\linewidth}
     \begin{tikzpicture}[scale=.5, every node={scale=.5}]
@@
       \begin{scope}
-        \clip (5.6,6.6) rectangle (9,11.5);
-        \capbrace{(5.6,6.5)}{(5.6,12)}{sp'}
+        \clip (5.1,6.6) rectangle (9,11.5);
+        \capbrace{(5.1,6.5)}{(5.1,12)}{sp'}
       \end{scope}
     \end{tikzpicture}
-  \caption{}
+  \caption{The stack after the trusted code has called the adversary. The
+    adversary's old stack pointer is still stored on the stack.}
   \label{fig:illu-adv-called}

### 2019-02-05 toplas/paper.tex
Place figure on its own page
Make various small improvements to the TOPLAS paper

@@
 
-Compromising software security is often based on attacks that break programming language properties relied upon by software authors, such as control-flow correctness, local-state encapsulation, etc.
+Compromising software security is often based on attacks that break programming language properties relied upon by software authors such as control-flow correctness, local-state encapsulation, etc.
 Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
 As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-mitigation arms race) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
-The latter techniques essentially emulate protection techniques on existing hardware, at the cost of performance, system complexity and/or security.
+The latter techniques essentially emulate protection techniques on existing hardware at the cost of performance, system complexity, and/or security.
 
@@
 remediate these limitations with a better security model at
-the hardware level. They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999},
-but have recently received renewed interest; in particular, the CHERI
+the hardware level. They are based on old
+ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that
+ have recently received renewed interest. in particular, the CHERI
 project has proposed new ideas and ways of tackling practical
@@
 Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}
-
 \begin{toplas}
@@
 \end{toplas}
-
 We have written a technical appendix~\citep{technical_appendix} which contains
@@
 We define register files $\reg$ and memories $\ms$ as functions mapping register names resp.\ addresses to words.
-The state of the entire machine is represented as a configuration that is either a running state $\Phi \in \ExecConfs$ containing a memory and a register file or a failed or halted state, where the latter is paired with the final state of memory.
+The state of the entire machine is represented as a configuration that is either a running state $\Phi \in \ExecConfs$ containing a memory and a register file or a failed or halted state where the latter is paired with the final state of memory.
 
@@
 cannot be used to read, write or execute and their address and range cannot be
-modified. They can only be used to jump to, but when that happens, their
+modified. They can only be used to jump to. When that happens, their
 permission changes to $\exec$. They can be used to represent a kind of closures:
@@
 contains local state (data or capabilities) and $[\addr,\addrend]$ contains
-instructions. The package is opaque to an adversary holding $c$ but when $c$ is
-jumped to, the instructions can start executing and have access to the local
+instructions. The package is opaque to an adversary holding $c$. When $c$ is
+jumped to however, the instructions can start executing and have access to the local
 data through the updated version of $c$, now in the $\pcreg$-register. 
@@
 if the capability has write authority for the requested location. However, the
-instruction is complicated by the presence of \emph{local capabilities}, modeled
+instruction is complicated by the presence of \emph{local capabilities} modeled
 after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
@@
 We point out that our local capabilities capture only a part of the semantics of
-local capabilities in CHERI. Specifically, in addition to the above, CHERI's
+local capabilities in CHERI. Specifically in addition to the above, CHERI's
 default implementation of the CCall exception handler forbids local capabilities
 from being passed across module boundaries. Such a restriction fundamentally
-breaks our calling convention, since we pass around local return pointers and
-stack capabilities. However, CHERI's CCall is not implemented in hardware, but
-in software, precisely to allow experimenting with alternative models like ours.
+breaks our calling convention since we pass around local return pointers and
+stack capabilities. However, CHERI's CCall is not implemented in hardware but
+in software precisely to allow experimenting with alternative models like ours.
 \dominique{Update to CHERI's new ccall story.}
@@
 The basic idea of our approach is simple: we stick to a single, rather standard,
-C stack and register-passed stack and return pointers, much like a standard C
+C stack and register-passed stack and return pointers much like a standard C
 calling convention. However, to prevent various ways of misusing this basic
@@
 returning to the adversary, (3) invoking the adversary, (4) returning from the
-adversary, (5) invoking an adversary callback and (6) having a callback invoked
+adversary, (5) invoking an adversary callback, and (6) having a callback invoked
 by the adversary.
@@
 prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
-the stack that the adversary has had access to, not just the parts that we have
+the stack that the adversary has had access to; not just the parts that we have
 used ourselves \itoplas{(\figurename~\ref{fig:illu-clear-all})}. Since we may be talking about a large part of memory, this
 requirement is the most problematic aspect of our calling convention for
-performance, but see \sectionname~\ref{sec:discussion} for how this might be
-mitigated.
-
-\begin{figure}[tb]
+performance (see the discussion in \sectionname~\ref{sec:discussion} for how this might be
+mitigated).
+
+\begin{figure}[ptb]
   \centering
@@
     \end{tikzpicture}
-    \caption{The stack just after we returned to the adversary and did take
+    \caption{The stack just after we returned to the adversary, and we did take
       care to clear our local stack frame.}
@@
     \end{tikzpicture}
-  \caption{The stack after an adversary has called us and we have called the adversary.}
+  \caption{The stack after an adversary has called us, and we have called the adversary.}
   \label{fig:illu-call-adv}
@@
 capability. If it were global, the adversary would be able to store away the
-return pointer in a global data structure (i.e.\ there exists a global capability for it), and jump to it later, in
+return pointer in a global data structure (i.e.\ there exists a global capability for it) and jump to it later in
 circumstances where this should not be possible. For example, they could store
 the return pointer, legally jump to it a first time, wait to be invoked again,
-and then jump to the old return pointer a second time, instead of the new return
+and then jump to the old return pointer a second time instead of the new return
 pointer received for the second invocation. Similarly, they could store the
 return pointer, invoke a function in our code, wait for us to invoke them again,
-and then jump to the old return pointer rather than the new one, received for the
+and then jump to the old return pointer rather than the new one received for the
 second invocation.
@@
     \end{tikzpicture}
-    \caption{The stack after the adversary has its stored the stack pointer in
+    \caption{The stack after the adversary has stored its stack pointer
       far up on the stack.}
@@
     \end{tikzpicture}
-  \caption{The stack after the adversary have called some trusted code.}
+  \caption{The stack after the adversary has called some trusted code.}
   \label{fig:illu-adv-call}
@@
   \caption{The stack after the trusted code has called the adversary. The
-    adversary's old stack pointer is still stored on the stack.}
+    adversary's old stack pointer is still available on the stack.}
   \label{fig:illu-adv-called}
@@
 
-\textbf{Returning from the adversary} So return pointers must be passed as local
+\textbf{Returning from the adversary} Return pointers must be passed as local
 capabilities. But what should their permissions be, what memory should they
@@
 
-Clearly, neither the old program counter or the old stack pointer should be accessible by the adversary. In
+Clearly, neither the old program counter nor the old stack pointer should be accessible by the adversary. In
 other words, the return pointer provided to the adversary must be a capability
@@
 over the entire activation record (containing the previous return and stack
-pointer), and its current address points to a number of restore instructions in
-the record, so that upon invocation, these instructions are executed and can
+pointer) and its current address points to a number of restore instructions in
+the record. Upon invocation, the instructions in the activation record are executed and can
 load the old stack pointer and program counter back into the register file. As
 the return pointer is an enter pointer, the adversary cannot get hold of the
-activation record's contents, but after invocation, its permission is updated to
+activation record's contents. However after invocation, its permission is updated to
 $\exec$, so the contents become available to the restore instructions.
@@
 machine's fine-grained memory protection should make it reasonably easy for a
-compiler to implement such protection, by making sure that only appropriately
+compiler to implement such protection by making sure that only appropriately
 bounded versions of the stack pointer are made available to source language
@@
 
-Our answer is that the adversary should provide it to us, just as we provide
+Our answer is that the adversary should provide it to us; just as we provide
 them with a stack pointer when we invoke their code. However, it is important
@@
 structure to access our local state. To prevent this from happening, we need to
-make sure the stack capability carries $\rwlx$ authority, since the system wide assumption then
+make sure the stack capability carries $\rwlx$ authority since the system wide assumption then
 tells us that the adversary cannot have global capabilities to our local stack.
@@
   \textbf{Modularity} The calling convention ensures well-bracketed calls and
-  local-state encapsulation for the caller, but not the callee. In the above
+  local-state encapsulation for the caller but not the callee. In the above
   presentation to make it easy to distinguish between the parties involved, we
@@
 Specifically, we employ a unary step-indexed Kripke logical relation with
-recursive worlds, and some additional special characteristics of our own.
-Step-indexing, Kripke logical relations and recursive worlds are techniques that
+recursive worlds and some additional special characteristics of our own.
+Step-indexing, Kripke logical relations, and recursive worlds are techniques that
 may be familiar from lambda calculus settings, but it may not be clear to the
 reader how they apply in this more low-level assembly language. Therefore, in the next section, we do
-not immediately dive into the details, but we first try to provide some informal
+not immediately dive into the details, but first we try to provide some informal
 intuition about how all of this machinery comes into play in our setting.
@@
 Note: even though the calling convention is the main application in this paper,
-the logical relation we construct is very general and should be regarded as an
+the logical relation we construct is very general and can be seen as a
+formulation of capability safety; hence it should be regarded as an
 independent contribution.
@@
 What differentiates a capability machine from a more standard assembly language
-is that we can bound the authority of an executing block of code, based solely
+is that we can bound the authority of an executing block of code based solely
 on the capabilities it has access to. Specifically, it does not matter which
-instructions are actually executed, i.e., the bound also applies to untrusted
+instructions are actually executed, i.e.\ the bound also applies to untrusted
 adversary code that has not been inspected or modified in any way.
@@
 But what does a ``bound on the authority'' of an executing block of code mean?
-In our setting, there are no externally observable side-effects and the only
+In our setting, there are no externally observable side-effects; the only
 primitive authority that code may hold is authority over memory. As such, the
@@
 our results will express authority bounds on code as \emph{safety with respect to such
-a world,} i.e., the fact that the code respects the invariants registered in
+a world,} i.e.\ the fact that the code respects the invariants registered in
 the world.
@@
 \paragraph{Safe values}
-So, let's say that we have a world $W$ expressing that the memory must contain
+Suppose we have a world $W$ expressing that the memory must contain
 value $42$ at address $0$, may contain arbitrary values at
@@
 should only include words that preserve $W$'s invariants under any interaction,
-but should otherwise be as liberal as possible. Numbers are clearly always safe,
+but should otherwise be as liberal as possible. Numbers are clearly always safe
 as they cannot be used to break invariants. Whether a capability is safe depends
 on the authority that it carries. In the above-described world, a read capability
-for address $0$ is safe, as it can only be used to read the value $42$, which is
+for address $0$ is safe as it can only be used to read the value $42$, which is
 itself safe. However, a write capability for address $0$ is not safe: it can be
-used to overwrite the memory at that address with a value other than $42$,
+used to overwrite the memory at that address with a value other than $42$
 breaking the invariant for that address.
@@
 More generally, we want to define that a read capability for memory range
-$[b,e]$ is safe, if the world guarantees that the words at those addresses are
-themselves safe. However, this definition is cyclic: what if the world
+$[b,e]$ is safe if the world guarantees that the words at those addresses are
+themselves safe. However, this definition is cyclic: suppose the world
 guarantees that the memory at address $a$ will contain a read capability for
-address $a$? The definition then just says that a read capability for address
+address $a$? Then the definition says that a read capability for address
 $a$ is safe if and only if the same read capability for address $a$ is safe. This form of
@@
 \paragraph{Future worlds}
-So worlds are defined as a set of invariants on the memory, but what if we
+Worlds are defined as a set of invariants on the memory, but what if we
 allocate fresh memory through malloc? We may want to establish new invariants
@@
 \paragraph{Invariants and Recursive Worlds}
-So worlds group a set of memory invariants, but how are they actually
+Worlds group a set of memory invariants, but how are they actually
 defined formally? We represent each invariant by a region
 $r \in \Regions$. We will see later that regions contain state
-machines to support a notion of evolvable invariants, but in every
-state, they also contain a predicate $H$ that defines which memory
-segments are acceptable in the current state of the invariant.
+machines to support a notion of evolvable invariants. Every
+state of the state machine contains a predicate $H$ that defines the valid
+memory segments.
 Unfortunately, it is not enough to just take
@@
 safe in the current world''. As explained, worlds may evolve, and the
-set of safe values may grow in future worlds, and therefore we need to
+set of safe values may grow in future worlds. Therefore, we need to
 index $H$ over worlds, i.e., take
-$H \in \Worlds \rightarrow \powerset{\MemSegments}$. We then end up
-with worlds containing regions with world-indexed predicates,
+$H \in \Worlds \rightarrow \powerset{\MemSegments}$. The result is worlds containing regions with world-indexed predicates,
 i.e., the set of worlds must be recursively defined. We will see how such a recursive
@@
 \paragraph{Local capabilities}
-The last thing we provide informal intuition about is how our results handle local capabilities and their special treatment by the hardware.
-Normally, when we invoke an untrusted piece of code and provide it with certain global capabilities, it may have stored those capabilities in memory and we will only be able to reinvoke the code if we can guarantee that those values are still valid.
-Formally, worlds represent the invariants that global capabilities' safety relies on and the reinvocation is only safe in future worlds, where the invariants are respected.
+When we invoke an untrusted piece of code and provide it with certain global capabilities, it may have stored those capabilities in memory.
+In this case, we will only be able to reinvoke the code if we can guarantee that those values are still valid.
+Formally, worlds represent the invariants that global capabilities' safety relies on and the reinvocation is only safe in future worlds where the invariants are respected.
 
@@
 The adversary has no way to store these local capabilities, so if we make sure that there are also no old local capabilities in the register file for the second invocation (including the capability being invoked), then the adversary cannot use them any more.%
-\footnote{We are ignoring write-local capabilities in this discussion.
-  If the adversary does have access to write-local capabilities in the first and second invocation, then the memory they address must entirely be cleared before the second invocation in order for the reinvocation to remain safe.}
-Therefore, we can allow the second invocation to happen in \emph{private} future worlds ($W' \futurestr W$), in which safe global capabilities remain safe, but local ones don't.
-This private future world relation is more liberal than the standard, public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain safe).
-Concretely, worlds may contain \emph{temporary} regions, representing invariants that only local capabilities may rely on for their safety, and which may be revoked (disabled) in private future worlds.
+\footnote{We ignore write-local capabilities in this discussion.
+  If the adversary does have access to write-local capabilities in the first and second invocation, then the memory they address must be cleared entirely before the second invocation in order for the reinvocation to remain safe.}
+Therefore, we can allow the second invocation to happen in any \emph{private} future worlds ($W' \futurestr W$) in which safe global capabilities remain safe but local capabilities do not.
+This private future world relation is more liberal than the standard public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain safe).
+Concretely, worlds may contain \emph{temporary} regions representing invariants that only local capabilities may rely on for their safety and that may be revoked (disabled) in private future worlds.
 
 Interestingly, this idea is a variant of a notion of public/private future worlds that has been previously used in the literature (see \sectionname~\ref{sec:discussion}).
-However, temporary regions are new in our setting and there is an interesting interplay with the recursiveness of the worlds: for a temporary region, the predicate $H \in \Worlds \rightarrow \powerset{\MemSegments}$ (which defines the safe memory segments in the current world) is only required to be monotone w.r.t.\ public future worlds (i.e.\ safe memory segments remain safe in public future worlds), but for permanent regions, it must be monotone w.r.t.\ private future worlds.
+However, temporary regions are new in our setting and there is an interesting interplay with the recursiveness of the worlds: for a temporary region, the predicate $H \in \Worlds \rightarrow \powerset{\MemSegments}$ (which defines the safe memory segments in the current world) is only required to be monotone w.r.t.\ public future worlds (i.e.\ safe memory segments remain safe in public future worlds).
+On the other hand for permanent regions, the world-indexed predicate must be monotone w.r.t.\ private future worlds.
 As a consequence, the memory for a permanent region may not contain local capabilities (as their safety would be broken in private future worlds), which in turn implies that only local capabilities may have write-local permission (a general sanity requirement when using local capabilities\footnote{In fact, local capabilities become useless as soon as the adversary has access to a single global, write-local capability.})\footnote{For explanation purposes, this discussion ignores certain ways to allow for local capabilities in a permanent region, for example, by not requiring that they are valid or requiring that they are local versions of valid global capabilities.}.
@@
 \label{subsec:worlds}
-\begin{figure}[htb]
+\begin{figure}[bt]
   \centering
@@
 can safely depend on the protocols imposed by temporary regions, but global
-capabilities cannot, since a global capability may outlive a temporary region
+capabilities cannot since a global capability may outlive a temporary region
 that is revoked. This is illustrated in \figurename~\ref{fig:cap-world}.
@@
 % future world of $W_1$. It is therefore safe to use the return pointer to return
-% to the adversary, since we have restored validity of any local state they might
+% to the adversary since we have restored validity of any local state they might
 % have stashed away.
@@
 The return capability would basically have to encapsulate this part of memory (perhaps indirectly) which means that from the moment we have been called, we will make sure there is only one stack available until our program returns from the initial call to it.}
-%p.10 "to prevent this from happening, we need to make sure the stack capability carries RWLX authority, since the system wide assumption then tells us that the adversary cannot have global capabilities to our local stack".
+%p.10 "to prevent this from happening, we need to make sure the stack capability carries RWLX authority since the system wide assumption then tells us that the adversary cannot have global capabilities to our local stack".
 It feels like the "L" capabitility is being overloaded to mean not just "this memory block has write-local authority and can contain local capabilities", but also to mean "this memory block is my stack and no other block is".
@@
 permission $\noperm$ are always in the value relation, irrespective of their
-address. Semantically, this makes perfect sense, since they are always safe
+address. Semantically, this makes perfect sense as they are always safe
 since they cannot be used for anything anyway. However, it also means we do not
@@
 
-For our purposes, this restriction is unproblematic, since we are only
+For our purposes, this restriction is unproblematic since we are only
 interested in proving semantic properties (e.g., an assertion will never fail).
@@
 \sectionname~\ref{sec:capab-mach-with} is a simple but representative
-model, modeled mainly after the
+model modeled mainly after the
 M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers

### 2019-02-05 toplas/preamble-illu.tex
Minor textual updates
Make various small improvements to the TOPLAS paper

@@
   \draw (-1.5,0) node {};
-  \draw (-1.5,15) node {};
+  \draw (-1.5,12) node {};
 }

### 2019-02-06 toplas/paper.tex
Update TOPLAS paper subsection about worlds

@@
 %As describe above, worlds are a collection of invariants. Formally, 
-
-A world is a
-finite map from region names, modeled as natural numbers, to regions that each
-correspond to an invariant of part of the memory.  We have three types of
-regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
-%$\perma$, $\temp$, and $\revoked$. The
-%$\perma$ and $\temp$ 
-Each permanent and temporary region contains a state transition system, with
-public and private transitions, to describe how the invariants are allowed to
-change over time. In other words, they are protocols for the region's
-memory. Protocols
-imposed by permanent regions stay in place indefinitely. Any capability, local
-or global, can depend on these protocols. Protocols imposed by temporary regions
-can be revoked in private future worlds. Doing this may break the safety of
-local capabilities but not global ones. This means that local capabilities
-can safely depend on the protocols imposed by temporary regions, but global
-capabilities cannot since a global capability may outlive a temporary region
-that is revoked. This is illustrated in \figurename~\ref{fig:cap-world}.
-
-For technical reasons, we do not actually remove a revoked temporary region from
-the world, but we turn it into a special revoked region that exists for this purpose.
-Such a revoked region contains no state transition system and puts no
-requirements on the memory. It simply serves as a mask for a revoked temporary
-region. Masking a region like this goes back to earlier work of \citet{Ahmed2004semantics} and was also used by \citet{Thamsborg:2011:KLR:2034773.2034831}.
+A world is a semantic model of the memory that carves out memories with a particular shape from all the possible memories.
+In our correctness proofs, worlds allow us to rely on the memory having a particular shape, but sometimes we will also have to guarantee that the memory has a certain shape.
+Essentially, a world is a collection of invariants. The memory satisfies the world when part of the memory satisfy each of the invariants.
+
+Worlds are represented as a finite map from region names, modeled as natural numbers, to regions that each correspond to an invariant on part of the memory.
+We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
+Each permanent and temporary region contains a state transition system with public and private transitions that describe how the invariant is allowed to change over time.
+In other words, they are protocols for the region's memory.
+Protocols imposed by permanent regions stay in place indefinitely.
+Any capability, local or global, can depend on these protocols.
+Protocols imposed by temporary regions can be revoked in private future worlds.
+Doing this may break the safety of local capabilities but not global ones.
+This means that local capabilities can safely depend on the protocols imposed by temporary regions, but global capabilities cannot since a global capability may outlive a temporary region that is revoked.
+This is illustrated in \figurename~\ref{fig:cap-world}.
+
+For technical reasons, we do not actually remove a revoked temporary region from the world, but we turn it into a special revoked region that exists for this purpose.
+Such a revoked region contains no state transition system and puts no requirements on the memory.
+It simply serves as a mask for a revoked temporary region.
+Masking a region like this goes back to earlier work of \citet{Ahmed2004semantics} and was also used by \citet{Thamsborg:2011:KLR:2034773.2034831}.
 
@@
 
-Regions are used to define safe memory segments, but this set may itself be world-dependent. In other
-words, our worlds are defined recursively. Recursive worlds are common
-in Kripke models and the following theorem uses the method of
-\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
-constructing them. The formulation of the lemma is technical, so we recommend
-that non-expert readers ignore the technicalities and accept that there exists a
-set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying
-the (recursive) equations in the theorem (where the $\blater$ operator can be
-safely ignored\footnote{The interested reader can find a brief coverage of \cofe{}'s
-and $\blater$ in the appendix.}).
+Regions are used to define safe memory segments, but this set may itself be world-dependent.
+In other words, our worlds are defined recursively.
+Recursive worlds are common in Kripke models and the following theorem uses the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for constructing them.
+The formulation of the lemma is technical, so we recommend that non-expert readers ignore the technicalities and accept that there exists a set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying the (recursive) equations in the theorem (where the $\blater$ operator can be safely ignored\footnote{The interested reader can find a brief coverage of \cofe{}'s and $\blater$ in Appendix~\ref{app:cofe}.}).
 % Short technical explanation
@@
 The future world relations model how memory may evolve over time.
-The \emph{public future world} $W' \futurewk W$ requires that $\dom(W') \supseteq
-\dom(W)$ and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$. That is, in a
-public future world, new regions may have been allocated, and existing regions
-may have evolved according to the public future region relation (defined below).
-The \emph{private future world} relation $W' \futurestr W$ is defined similarly,
-using a private future region relation. The \emph{public future} region relation is the simplest. It satisfies
+We have a \emph{public} and a \emph{private} future world relation that respectively what changes any capability can rely on and what changes only temporary regions can rely on.
+As temporary capabilities fall within the category of all capabilities, the public future relation is subsumed in the private future relation.
+
+The public future world $W' \futurewk W$ requires that $\dom(W') \supseteq \dom(W)$ and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$.
+That is in a public future world, new regions may have been allocated, and existing regions may have evolved according to the public future region relation (defined below).
+The private future world relation $W' \futurestr W$ is defined similarly, using a private future region relation.
+The public future region relation is the simplest.
+It satisfies
 % \dominique{is defined by?}
@@
 \end{mathpar}
-Both temporary and permanent regions are only allowed to transition according to
-the public part of their transition system. Additionally, revoked regions must
-either remain revoked or be replaced by a temporary region. This means that the
-public future world relations allows us to reinstate a region that has been revoked
-earlier. The \emph{private future region} relation satisfies:
+In public future worlds, both temporary and permanent regions are only allowed to transition according to the public part of their transition system.
+Additionally, revoked regions must either remain revoked or be replaced by a temporary region.
+This means that the public future world relations allows us to reinstate a region that has been revoked earlier.
+The private future region relation satisfies:
 \begin{mathpar}
@@
 \end{mathpar}
-Here, revocation of temporary regions is allowed. In fact, temporary regions can
-be replaced by an arbitrary other region, not just the special $\revoked$.
-Conversely, $\revoked$ regions may also be replaced by any other region. On
-the other hand, permanent regions cannot be masked away. They are only allowed
-to transition according to the private part of the transition system.
+Here, revocation of temporary regions is allowed.
+In fact, temporary regions can be replaced by an arbitrary region not just the special $\revoked$.
+Conversely, $\revoked$ regions may also be replaced by any other region.
+On the other hand, permanent regions cannot be masked away.
+They are only allowed to transition according to the private part of the transition system.
 
@@
   This is why we allow revoked regions to be replaced by other regions in the two future world relations.
-  This may seem unnecessary as the future world relation allows new regions to be added, but in private future worlds, for technical reasons, we need to be able to reinstate old regions with the same region names as they had in a past world.
+  This may seem unnecessary as the future world relation allows new regions to be added.
+However, for technical reasons in private future worlds, we need to be able to reinstate old regions with the same region names as they had in a past world.
   Further, we do not want public future worlds to be able to prevent reinstation in private future worlds.
-  Private future worlds cannot revoke permanent regions, so we cannot allow public future worlds to us permanent regions as a mask for a revoked region.
+  Private future worlds cannot revoke permanent regions, so we cannot allow public future worlds to use permanent regions as masks for revoked regions.
 \end{toplas}
 
-Notice that the public future region relation is a subset of the private future
-region relation.
 %domi: the below seems hard to understand at this moment.

### 2019-02-06 toplas/paper.tex
Update TOPLAS LR section with new read cond description

@@
 The logical relation defines semantically when values, program
-counters, and configurations are capability safe. The definition is
-found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds} and
+counters, and configurations are capability safe. The logical relation is defined
+found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and
 we provide some explanations in the following paragraphs. For space
@@
 In fact, the machine often resorts to failing when an unauthorized access is
-attempted, such as loading from a capability without read permission. This is
+attempted such as loading from a capability without read permission. This is
 similar to \citet{Devriese:2016ObjCap}'s logical relation for an untyped
-language, but unlike typical logical relations for typed languages, which
-require that programs do not fail.
-
-The \emph{register-file relation} $\stdrr$ defines safe register-files as those
-that contain safe words (i.e.\ words in the $\stdvr$-relation defined below) in all registers but $\pcreg$.
-The \emph{expression relation} $\stder$ defines that a word is safe to use as a
-program counter if it can be plugged into a safe register file (i.e.\ a register
-file in $\stdrr$) and paired with a memory satisfying the world to become a safe
-configuration. Note that integers and non-executable capabilities (e.g.
-$\readonly$ and $\entry$ capabilities) are considered safe program counters
-because when they are plugged into a register file and paired with a memory, the
-execution will immediately fail, which is safe.
+language but unlike typical logical relations for typed languages that
+require programs to not fail.
+
+The \emph{register-file relation} $\stdrr$ defines safe register-files as those that contain safe words (i.e.\ words in the $\stdvr$-relation defined below) in all registers but the $\pcreg$-register.
+The \emph{expression relation} $\stder$ defines what words are safe to use as a program counter.
+A word is safe to use as a program counter when it can be used to form a safe configuration by plugging it into the $\pcreg$-register of a safe register file (i.e.\ a register file in $\stdrr$) and pairing it with a memory satisfying the world.
+Note that integers and non-executable capabilities (e.g.\ $\readonly$ and $\entry$ capabilities) are considered safe program counters because when they are plugged into a register file and paired with a memory, the execution will immediately fail, which is safe.
 
@@
 \itoplassug{Explain why we use $\UPred{}$ with reference to the introduction of this section.}
-We make the
-value relation as liberal as possible by considering what is the most we can
-allow an adversary to use a capability for without breaking the memory
-protocols. Non-capability data is always safe because it provides no
-authority. Capabilities give the authority to manipulate memory and potentially
-break memory protocols, so they need to satisfy certain conditions to be
-safe. In \figurename~\ref{fig:perm-conds}, we define such a condition for each
-kind of permission a capability can have.
-
-For capabilities with read permission, the $\readCond{}$ ensures that it can
-only be used to read safe words, i.e.\ words in the value relation. To guarantee
-this, we require that the addressed memory is governed by a region $W(r)$ that
-imposes safety as a requirement on the values contained. This safety requirement
-is formulated in terms of a standard region $\iota^\pwl_{[\start,\addrend]}$. It simply
-requires all the words in the range $[\start,\addrend]$ to be safe, i.e.\ in the
-value relation. Requiring that $W(r)\nsubsim[n]\iota^\pwl_{[\start,\addrend]}$
-means that $W(r)$ must accept only memory segments containing safe values like
-$\iota^\pwl_{[\start,\addrend]}$, but can be even more restrictive if needed
-(the $\nsubsim[n]$-relation is defined in Appendix~\ref{app:logical-relation}). The
-read condition also takes into account the locality of the capability because,
-generally speaking, global capabilities should only depend on permanent regions.
-Concretely, we use the function $\var{localityReg}(\gl,W)$, which projects out
-all active (non-revoked) regions when the locality $\gl$ is local, but only the
-permanent regions when $\gl$ is global. The definition of the standard region
-$\iota^\pwl_{[\start,\addrend]}$ can be found
-in \figurename~\ref{fig:std-reg}; it makes use of the isomorphism $\xi$ from
-Theorem~\ref{thm:world-existence}.
+We make the value relation as liberal as possible by defining it based on the principle ``what is the most we can allow an adversary to use a capability for without breaking the memory protocols.''
+
+Non-capability data is always safe because it provides no authority.
+Capabilities give the authority to manipulate memory and potentially break memory protocols, so they need to satisfy certain conditions to be safe.
+In \figurename~\ref{fig:perm-conds}, we define such a condition for each kind of permission a capability can have.
+
+Capabilities with read permission cannot directly break memory protocols because they cannot make changes to the memory.
+However, a read capability could be used to read a capability that can break memory protocols.
+For this reason, a capability with read permissions is only safe when it can be used only to read safe capabilities.
+This is captured by the $\readCond{}$-condition. 
+More precisely, the condition requires the memory addressed by the capability to actually be governed by some region $W(r)$.
+This region should at least require all the memories accepted by the state interpretation function to contain safe words but stricter requirements may be imposed on the memories.
+This is expressed in terms of an upper bound on the possibly permitted memories defined in terms of a so called ``standard region''.
+In this case, the standard region is $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
+The standard region only accept memories with address range $[\start,\addrend]$.
+More importantly, the standard region's state interpretation function only accepts memory segments that only contain safe words.
+The relation the actual region $W(r)$ and the upper bound $\iota^\pwl_{[\start,\addrend]}$ is $\iota^\pwl_{[\start,\addrend]}$ (defined in Appendix~\ref{app:logical-relation}).
+Essentially, the relation requires the two regions state transition systems to be the same.
+Further in the current state of region $W(r)$ and in any world, the state interpretation function should only allow memories that are also allowed in the standard region.
+
+The read condition also takes the locality of a read capability into account because, generally speaking, generally speaking, global capabilities should only depend on permanent regions.
+In $\readCond{}$, this is manifested by $\var{localityReg}(\gl,W)$ which projects out all active (non-revoked) regions when the locality $\gl$ is local, but only the permanent regions when $\gl$ is global.
 

### 2019-02-07 toplas/paper.tex
Rewrite TOPLAS paragraphs on read and write condition

@@
 \newcommand{\scall}[3]{\mathtt{scall} \; #1([#2],[#3])}
+
+% Text
+\newcommand{\tpwl}{permit-write-local standard region}
+\newcommand{\tnwl}{no-write-local standard region}
 
@@
 This is expressed in terms of an upper bound on the possibly permitted memories defined in terms of a so called ``standard region''.
-In this case, the standard region is $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
-The standard region only accept memories with address range $[\start,\addrend]$.
-More importantly, the standard region's state interpretation function only accepts memory segments that only contain safe words.
-The relation the actual region $W(r)$ and the upper bound $\iota^\pwl_{[\start,\addrend]}$ is $\iota^\pwl_{[\start,\addrend]}$ (defined in Appendix~\ref{app:logical-relation}).
+In this case, it is a \tpwl{} $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
+The \tpwl{} only accept memories with address range $[\start,\addrend]$.
+More importantly, the \tpwl{}'s state interpretation function only accepts memory segments that only contain safe words.
+The relation between the actual region in the world and the upper bound is $W(r) \nsubsim[n] \iota^\pwl_{[\start,\addrend]}$ (defined in Appendix~\ref{app:logical-relation}).
 Essentially, the relation requires the two regions state transition systems to be the same.
@@
 
-The read condition also takes the locality of a read capability into account because, generally speaking, generally speaking, global capabilities should only depend on permanent regions.
-In $\readCond{}$, this is manifested by $\var{localityReg}(\gl,W)$ which projects out all active (non-revoked) regions when the locality $\gl$ is local, but only the permanent regions when $\gl$ is global.
-
-For a capability with write permission, $\writeCond{}$ must be satisfied for the
-capability's range of authority. An adversary can use such a capability to write
-any word they can get a hold of, and we can safely assume that they can only get
-a hold of safe words, so the region governing the relevant memory must allow any
-safe word to be written there. In order to make the logical relation as liberal
-as possible, we make this a lower bound of what the region may allow. 
-% LB: omitted the following for brevity:
-% From another point of view, this lower bound is also an upper bound on the amount of
-% trust that can be placed in the contents of memory that an adversary has been
-% given write access to. 
-For write capabilities, we also have to take into account
-the two flavours of write permissions: write and write-local. In the case of
-write-local capabilities, the region needs to allow (at least) any safe word to
-be written, but in the case of write capabilities, the capability cannot be used
-to write local capabilities, so the region only needs to allow safe non-local
-values. In the write condition, this is handled by parameterizing it with a
-region. For the write-local capabilities the write condition is applied with the
-standard region $\iota^\pwl_{[\start,\addrend]}$ that we described previously. For
-the write capabilities we use a different standard region
-$\iota^\nwl_{[\start,\addrend]}$ which requires that the words in
-$[\start,\addrend]$ are non-local and safe. As before, we use
-$\var{localityReg}$ to pick an appropriate region based on the capability's
-locality. Finally, there is a technical requirement that the region must be
-\emph{address-stratified}. Intuitively, this means that if a region accepts two
-memory segments, then it must also accept every memory segment ``in between'',
-that is every memory segment where each address contains a value from one of the
-two accepted memory segments. An interesting property of the write condition is
-that it prohibits global write-local capabilities which, as discussed in
-\sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of
-local capabilities.
-The standard region $\iota^\nwl_{[\start,\addrend]}$ is defined in \figurename~\ref{fig:std-reg}.
+The locality of capabilities play a role in whether or not they are safe.
+Generally speaking, global capabilities should only depend on permanent regions.
+This is expressed in $\readCond{}$ with $\var{localityReg}(\gl,W)$ which projects out all the regions the capability can depend on based on its locality.
+Specifically, if the capability is local, then all active regions (non-revoked) are project.
+On the other hand, if the capability is global, then only the permanent regions are projected.
+
+Capabilities with write permission can be used to write to memory, so to define when a capability with write permission is safe we ask ourselves what an adversary should be allowed to do with it.
+An adversary should at least be able to write safe words to memory.
+Safe words cannot be used to break memory invariants, so we will permit any safe word to be written.
+A capability with write permission\footnote{On the capability machine we consider, write permission always comes with read permission, so we could have made a stricter condition for write capabilities.}
+only allows you to write to memory not read from it, so we can allow anything to be written to memory as long as it cannot be read back and used to break memory protocols.
+For this reason, the condition on capabilities with write-permission is defined as a lower bound on what can be written where the lower bound is ``any safe word''.
+
+The condition on write capabilities is complicated by the fact that we have two flavours of write permission: write and write-local.
+A write-local capability can be used to write both local and global capabilities, so it is (at least) allowed to write any safe capabilities.
+Write capabilities, on the other hand, cannot write local capabilities, so we set the lower-bound as non-local safe words.
+
+The requirements on write-capabilities are captured by the $\writeCond{}$-condition, and it is defined in a similar fashion to the $\readCond{}$-condition.
+The regions the write-capability may depend on are projected from the world with the $\var{localityReg}$-function, and $\nsupsim[n-1]$ is used to relate the actual region to the lower bound.
+We need a different lower-bound depending on whether the permission of the capability is write or write-local, so $\writeCond{}$ is parameterized with the region it actually uses.
+If the capability has write-local permission, then we use the \tpwl{}.
+If capability only has write permission, then we use a \tnwl{} $\iota^\nwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}).
+In addition to the requirements of a \tpwl, the \tnwl{} requires all words in the accepted memory segments to be non-local.
+
+Finally, there is a technical requirement that the region must be \emph{address-stratified}.
+Intuitively, this means that if a region accepts two memory segments, then it must also accept every memory segment ``in between'' in the sense that it should be possible to come from one memory segment to the other.
+Our capability machine can only update one memory address at a time, so it should be accepted to construct a memory that is on its way to become the one we want to accept.
+Specifically, if two memory segments are accepted by a region, then the region must also accept every memory segment where each address contains a value from one of the two accepted memory segments.
+
+Due to the permission combinations on the capability machine we consider, a $\writeCond{}$ always comes with a $\readCond{}$ which creates a somewhat tight bound for the possible regions that make a capability safe.
+Further, we could have made the $\writeCond{}$ based on the fact that a writer permission always comes with a read permission, but we opted to make the logical relation as general as possible, so it can be reused in a setting with a richer permission hierarchy.
+An interesting property of the write condition is that it prohibits global write-local capabilities which, as discussed in \sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of local capabilities.
 

### 2019-02-08 toplas/paper.tex
Resolve a TODO.

@@
 regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
-assume that memory allocated through malloc cannot be freed.
+assume that memory allocated through malloc cannot be freed.\footnote{In more realistic settings, reusing freed memory on a capability machine can be made safe by checking or enforcing that there are no dangling pointers to the freed memory, as implemented, for example, in CHERI-JNI~\citep{chisnall_cheri_2017}.
+}
 

### 2019-02-10 toplas/paper.tex
Fix layout of a figure
Not sure why this is needed...
Fix a figure reference and one of the pictures in Fig.4
Rewrite condition paragraph in TOPLAS paper

--- 2019-02-07 toplas/paper.tex
@@
 Further, we could have made the $\writeCond{}$ based on the fact that a writer permission always comes with a read permission, but we opted to make the logical relation as general as possible, so it can be reused in a setting with a richer permission hierarchy.
-An interesting property of the write condition is that it prohibits global write-local capabilities which, as discussed in \sectionname~\ref{sec:stack-and-return-pointer}, is necessary for any safe use of local capabilities.
 
@@
 In contrast, enter capabilities are opaque and can only be used to jump to the
-address they point to.%
+address they point to.
+%
 \begin{toplas}%
-  This is why the $\entryCond{}$ depends on the current address of the capability, unlike for other types of capabilities.
+  This is why $\entryCond{}$ depends on the current address of the capability unlike for other types of capabilities.
 \end{toplas}
@@
 revoked). For local capabilities, it suffices to be safe in \emph{public} future
-worlds, where temporary regions are still present.
+worlds where temporary regions are still present.
 
@@
   In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
-  This would indicate that our logical relation should imply the same, namely that capabilities with write-local permission are local:
+  This indicates that our logical relation should imply the same, namely that capabilities with write-local permission are local.
   \begin{lemma}[Stack capabilities are local]~

### 2019-02-10 toplas/paper.tex
[TOPLAS] Update capability safety section

@@
 
-\subsection{Safety of the Capability Machine}
+\subsection{Capability Machine Safety}
 With the logical relation defined, we can now state the fundamental theorem of
@@
 
-    The rest of the proof considers the different ways the step $\Phi \step
-    \Phi'$ could have occurred, depending on the instruction being executed. For
-    each of these cases, we argue (1) that $\Phi'$ is consistent with the world
+    The rest of the proof considers the different possibilities for the step $\Phi \step
+    \Phi'$, i.e. it considers each of the instructions that could have been executed. For
+    each of these cases, we argue that (1) $\Phi'$ is consistent with the world
     in the sense that the register-file and memory still respect the world and
-    (2) that the rest of the execution respects the world. Depending on where
+    that (2) the rest of the execution respects the world. Depending on where
     the $\pcreg$ in $\Phi'$ comes from, the second result is proven in one of
     two ways. If the step to $\Phi'$ was a jump, then the new $\pcreg$ is one of
-    the safe values in $\Phi$'s registers and the value relation can be used to
-    argue that those can be safely jumped to. On the other hand, if the $\pcreg$
+    the safe values in $\Phi$'s registers, and the value relation is used to
+    argue that the jump is safe. On the other hand, if the $\pcreg$
     was just incremented, then we can apply the induction hypothesis.
 
-    In order to argue (1), we consider what configurations the operational
-    semantics allows us to get to from the initial state. If we consider the
-    memory in $\Phi'$, then it either (a) remains unchanged or (b) one address
-    has been updated and in that case, the register-file contains an appropriate
-    capability for writing. The latter occurs when the instruction being
-    executed is a $\texttt{store}$, and otherwise the former. In case (a), we
-    can conclude that the memory still respects the world just by downwards
-    closure of memory satisfaction. In case (b), we use an auxiliary lemma that
-    uses the safety of the write capability used by the $\texttt{store}$
-    instruction to show that the updated memory satisfies the world. To show
-    that the updated register-file is safe, we consider the changes made to it
-    by all instructions in separate lemmas and show that they all preserve
-    safety of the register file.
+    In order to argue (1), we look at the possible states for $\Phi$ according to the operational semantics.
+    If we consider the memory in $\Phi'$, then it either (a) remains unchanged or (b) one address has been updated and the register-file contains an appropriate capability for writing.
+    The latter occurs when the executed instruction is $\texttt{store}$.
+    Otherwise we are in the former case.
+    In case (a), the memory hasn't changed, so we conclude that the memory still respects the world simply by downwards closure of memory satisfaction.
+    In case (b), we use an auxiliary lemma that uses the safety of the write capability used by the $\texttt{store}$ instruction to show that the updated memory satisfies the world.
+    To show that the updated register-file is safe, we consider the changes made to it by all instructions in separate lemmas and show that they all preserve safety of the register file.
 
@@
 
-The Fundamental Theorem can be understood as a general expression of the
-guarantees offered by the capability machine, an instance of a general property
-called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}. To
-understand this, consider that the theorem says the capability
-$((\perm,\gl),\start,\addrend,\addr)$ is safe as a program counter, without any
-assumption about what instructions it actually points to (the only assumptions
-we have are about the read or write authority that it carries). As such, the
-theorem expresses the capability safety of the machine, which guarantees that
-\emph{any} instruction is fine and will not be able to go beyond the authority
-of the values it has access to. We demonstrate this in
-\sectionname~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about
-capabilities that point to arbitrary instructions. The relation between
-Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow
-correctness, will also be shown by example in \sectionname~\ref{sec:examples} as the
-examples depend on these properties for correctness.
+The Fundamental Theorem can be understood as a general expression of the guarantees offered by the capability machine which is an instance of a general property called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}.
+The theorem says that an arbitrary capability $((\perm,\gl),\start,\addrend,\addr)$ is safe as a program counter without making any assumptions about what program it points to (the only assumptions we have are about the read or write authority that it carries).
+As such, the theorem expresses the capability safety of the machine which guarantees that \emph{any} instruction is safe to execute and will not be able to go beyond the authority of the values it has access to.
+We demonstrate this in \sectionname~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about capabilities that point to arbitrary instructions.
+The relation between Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow correctness will also be shown by example in \sectionname~\ref{sec:examples} as the examples depend on these properties for correctness.
+
 % See the technical appendix~\citep{technical_appendix} for a detailed

### 2019-02-10 toplas/paper.tex
[TOPLAS] Update malloc specification secion

@@
   \label{sec:malloc}
-  In the examples we present in \sectionname~\ref{sec:examples}, we will assume
-  the existence of a trusted malloc routine so that both the trusted code and
-  the adversary can allocate new memory. Malloc is considered part of the
-  trusted computing base, as mentioned in
-  \sectionname~\ref{sec:capab-mach-with}. This is unavoidable: if we cannot
-  trust malloc, then we cannot use the memory it allocates as we have no idea
-  whether it is aliased by some untrusted program.
-
-  Our semantic model is not specific to a particular implementation of malloc,
-  so rather than providing the implementation, we provide a specification that
-  malloc must satisfy. The specification expresses standard expectations of how
-  malloc should behave, but making it realistic requires using some of the
-  technical machinery from the logical relation. As such, this section is a bit
-  technical and can safely be skipped on first read. The malloc specification is
-  presented in \figurename~\ref{fig:malloc-spec}, and in the following, we will
-  connect the description of the specification by referring to the item
-  described.
-
-We require a global capability for
-invoking malloc (because if the capability were local, then a program with
-access to malloc would have to give up this access when invoking untrusted
-code, \ref{item:malloc-spec:eg}). The capability is assumed to have enter
-permission (\ref{item:malloc-spec:eg}), so that malloc can
-protect its internal state even when the capability is shared with untrusted
-code.
-
-In addition to these syntactic requirements, we also specify standard
-expectations for how malloc behaves. Intuitively, we require that when malloc is
-invoked with a length argument, then it will return a capability for a fresh
-piece of memory of that size. It should be fresh in the sense that malloc has
-not already given out a capability for any part of that memory before and will
-not do so in the future. Also, when invoked with a nonsensical length argument
-such as a negative integer or a capability, malloc should simply fail. However,
-formulating these requirements is harder than one might expect. The problem is
-that a realistic implementation of malloc needs to rely on internal state that
-changes after every invocation and relies on invariants on that state. We can
-only expect that malloc behaves according to its specification, if its internal
-state satisfies the current state of the invariants in an executing system. To
-express this, we will use the semantic model defined in
-Section~\ref{sec:logical-relation}.
-
-To allow malloc implementations to rely on internal state and invariants for
-that state, we assume an initial region for malloc (\ref{item:malloc-spec:reg}). The region is assumed to be
-permanent (since safety of the global malloc capability will depend on its
-presence, \ref{item:malloc-spec:perm}).
-% DOMI: I don't understand what the following uncommented text means
-% This also means that the malloc specification should guarantee that
-% use of malloc is safe at any point of execution no matter who has invoked it. To
-% this end, the requirement on the initial malloc region should also hold true in
-% any private future world.
-% priv future means that malloc should not depend on local capabilities?
-Furthermore, we want to express that malloc does not depend on any other memory
-than its own internal state. This is expressed by a restriction on the malloc
-region's state interpretation function, which (as explained in
-\sectionname~\ref{subsec:worlds}) defines what memory segments it permits in a
-given world. We require that the accepted memory segments only depend on the
-presence of the malloc region itself, i.e.\ that in any world, the same memory
-segments are accepted if we remove all regions except the malloc region. This
-property should continue to hold throughout execution, so it must hold true for
-any private future region of the initial malloc region (\ref{item:malloc-spec:independent}).
-
-The malloc specification then dictates what malloc should do when invoked in a
-memory with its internal state valid according to the malloc region (some future
-evolution of the initial malloc region). If malloc is invoked with an invalid
-length argument (that is, a negative integer or a capability), then we simply
-require malloc to fail (\ref{item:malloc-spec:fail}). This part of the
-specification does not actually rely on the malloc region: for simplicity, we
-assume malloc does not need its internal state to check the argument. When
-malloc is invoked with a valid length (\ref{item:malloc-spec:correct-exec}),
-then it should return a fresh memory segment of the correct length. This segment
-is required to come from the footprint of malloc, i.e.\ the memory owned by the
-malloc region before the call. After malloc returns, we require the malloc
-region to have evolved (according to the public future world relation) to a new
-state where the new footprint is disjoint from the allocated memory. This
-implies that future invocations of malloc can never return previously-allocated
-memory.
-
-For convenience, we also require that malloc returns the non-argument registers
-of the register-file unchanged after the call. This allows the caller to keep
-private capabilities in the register file, without having to protect them by
-storing them in a private stack frame.
-
-\begin{figure}[htb]
-  \centering
+  In the examples presented in \sectionname~\ref{sec:examples}, we will assume the existence of a trusted malloc routine, so that both the trusted code and the adversary are able to allocate new memory.
+  Malloc is considered part of the trusted computing base as mentioned in \sectionname~\ref{sec:capab-mach-with}.
+  This is unavoidable: if we cannot trust malloc, then we cannot use the memory it allocates as we have no idea whether it is aliased by some untrusted program.
+
+  Our semantic model is not specific to a particular implementation of malloc, so rather than providing the implementation we provide a malloc specification.
+  The specification expresses standard expectations about the behaviour of malloc, but making the specification realistic requires some of the technical machinery from the logical relation.
+  As such, this section is a bit technical and can safely be skipped on first read.
+  Definition~\ref{spec:malloc} is the malloc specification.
+  Following the definition, we provide an informal description of the definition with references to each of the items in the definition.
+
 \begin{definition}[Malloc Specification]
@@
 
-  \caption{The specification of malloc.}
-  \label{fig:malloc-spec}
-\end{figure}
+  We require a global capability for invoking malloc (because if the capability were local, then a program with access to malloc would have to give up this access when invoking untrusted code, \ref{item:malloc-spec:eg}).
+  The capability is assumed to have enter permission (\ref{item:malloc-spec:eg}), so malloc can protect its internal state even when the capability is shared with untrusted code.
+
+  In addition to these syntactic requirements, we also specify standard behavioural expectations of malloc.
+  Intuitively, we require that when malloc is invoked with a length argument, then it will return a capability for a fresh piece of memory of that size.
+  It should be fresh in the sense that malloc has not already allocated any of that memory and will not do so in the future.
+  Also, when invoked with a nonsensical length argument such as a negative integer or a capability, malloc should simply fail.
+  However, formulating these requirements is harder than one might expect.
+  The problem is that a realistic implementation of malloc needs to rely on internal state that changes after every invocation and relies on invariants on that state.
+  We can only expect that malloc behaves according to its specification if its internal state satisfies the current state of the invariants in an executing system.
+  We express this in terms of the semantic model from \sectionname~\ref{sec:logical-relation}.
+
+To allow malloc implementations to rely on internal state and invariants for
+that state, we assume an initial region for malloc (\ref{item:malloc-spec:reg}). The region is assumed to be
+permanent (since safety of the global malloc capability will depend on its
+presence, \ref{item:malloc-spec:perm}).
+% DOMI: I don't understand what the following uncommented text means
+% This also means that the malloc specification should guarantee that
+% use of malloc is safe at any point of execution no matter who has invoked it. To
+% this end, the requirement on the initial malloc region should also hold true in
+% any private future world.
+% priv future means that malloc should not depend on local capabilities?
+Furthermore, we want to express that malloc does not depend on any other memory
+than its own internal state. This is expressed by a restriction on the malloc
+region's state interpretation function which (as explained in
+\sectionname~\ref{subsec:worlds}) defines what memory segments it permits in a
+given world. We require that the accepted memory segments only depend on the
+presence of the malloc region itself, i.e.\ in any world the same memory
+segments are accepted if we remove all regions except the malloc region. This
+property should continue to hold throughout execution, so it must hold true for
+any private future region of the initial malloc region (\ref{item:malloc-spec:independent}).
+
+The malloc specification also dictates what malloc should do when invoked in a
+memory with its internal state valid according to the malloc region (some future
+evolution of the initial malloc region). If malloc is invoked with an invalid
+length argument (that is, a negative integer or a capability), then we simply
+require malloc to fail (\ref{item:malloc-spec:fail}). This part of the
+specification does not actually rely on the malloc region: for simplicity, we
+assume malloc does not need its internal state to check the argument. When
+malloc is invoked with a valid length (\ref{item:malloc-spec:correct-exec}),
+then it should return a fresh memory segment of the correct length. This segment
+is required to come from the footprint of malloc, i.e.\ the memory owned by the
+malloc region before the call. After malloc returns, we require the malloc
+region to have evolved (according to the public future world relation) to a new
+state where the new footprint is disjoint from the allocated memory. This
+implies that future invocations of malloc can never return previously-allocated
+memory.
+
+For convenience, we also require that malloc returns the non-argument registers
+of the register-file unchanged after the call. This allows the caller to keep
+private capabilities in the register file, without having to protect them by
+storing them in a private stack frame.
+
 As described previously, the specification of malloc ensures that malloc has no
-capabilities pointing \emph{out} of malloc. It does not, however, say anything about
-capabilities that point \emph{in} to malloc. We obviously cannot
-allow this if we want to trust malloc. We have chosen to keep the malloc
-specification simple and let this assumption be in the lemmas that use malloc.
-It is sufficient for these lemmas to require that there are no outside
-capabilities for malloc in the initial configuration as capabilities cannot
-appear out of thin air and the malloc specification makes sure that capabilities
-are not leaked.
-
-Malloc should not just be available to trusted programs, but also to possibly
+capabilities pointing \textit{out} of malloc. It does not, however, say anything
+about capabilities that point \textit{in} to malloc. If we want to be able to
+trust malloc, we obviously cannot allow arbitrary capabilities to point in to
+it. We have chosen to keep the malloc specification simple and let this
+assumption be in the lemmas that use malloc. It is sufficient for these lemmas
+to require that there are no outside capabilities for malloc in the initial
+configuration as capabilities cannot appear out of thin air, and the malloc
+specification makes sure that capabilities are not leaked.
+
+Malloc should not just be available to trusted programs but also to possibly
 malicious programs. This is safe as it follows from the specification that the malloc

### 2019-02-10 toplas/paper.tex
[TOPLAS] Update the macro section

@@
 convention to enforce these properties. However, the programs that
-run on our capability machine are assembly programs and even program examples that would
+run on our capability machine are assembly programs. Program examples that would
 be small in a high-level language become big and unintelligible at this
@@
 We define a number of reusable macros capturing the calling convention as well
-as other conveniences. The macros that utilise the stack assume that it is
+as other conveniences. The macros that utilize the stack assume that it is
 available in register $r_\stk$.
@@
 The macro \texttt{scall
-    $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures
-those parts of the calling convention related to actually transferring control
-to some adversarial code. Specifically, it pushes the contents of the private
-registers, $\overline{r_{\var{priv}}}$, to the stack. Then it pushes the
-``restoration'' code to the stack. The restoration code will be executed as the
-first thing upon return, and restores the stack pointer and the old program
-counter. After the restoration code is pushed to the stack, a copy of the pc is
-pushed to the stack after it has been adjusted to point to the first instruction
-after the jump. From the stack pointer, a protected return pointer is created by
-adjusting it to point to the return code and encapsulating it by restricting its
-permission to $\entry$. Next, the stack pointer is restricted to the unused
-part, and the unused part of the stack is cleared (as discussed in
-Section~\ref{sec:stack-and-return-pointer}). Finally, the non-argument registers
-are cleared and we jump to $r$. Upon return after the restoration code has been
-executed, the restoration code is popped from the stack and the private words we
-pushed to the stack before the call are popped to the private registers. The
+  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures the
+parts of the calling convention related to actually transferring control to
+adversarial code. Specifically, it pushes the contents of the private registers,
+$\overline{r_{\var{priv}}}$, to the stack and pushes the ``restoration'' code to
+the stack. The restoration code is executed as the first thing upon return; it
+restores the stack pointer and the old program counter. After the restoration
+code is pushed to the stack, \texttt{scall
+  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} adjusts the pc
+to point to the first instruction after the jump and pushes it to the stack. The
+\texttt{scall}macro constructs a protected return pointer from a stack pointer by
+adjusting it to point to the first instruction of the return code and
+encapsulating it by restricting its permission to $\entry$. Next, the \texttt{scall}
+macro reduces the range of authority of the stack pointer to the unused part of
+the stack and clears it(as discussed in
+Section~\ref{sec:stack-and-return-pointer}). Finally, the \texttt{scall} macro clears
+non-argument registers and jumps to $r$. Upon return after the restoration code
+on the stack has been executed, the \texttt{scall} macro pops the restoration code from
+the stack and restores the private state by popping it from the stack to the
+appropriate registers. \figurename~\ref{fig:scall-impl} displays the
 implementation of \texttt{scall} and the restoration code used by
-\texttt{scall} are presented in \figurename~\ref{fig:scall-impl}. The implementation of \texttt{scall} uses
-some of the macros we present next.
+\texttt{scall}. The implementation of \texttt{scall} uses some of the macros we
+present next.
 
@@
     \begin{minipage}[t]{.29\linewidth}
-\begin{lstlisting}[title=\footnotesize{The restoration code used in \texttt{scall}.}, captionpos=b,xleftmargin=2em,frame=single,framexleftmargin=1.5em]
+\begin{lstlisting}[title=\footnotesize{The restoration code used in \texttt{scall}.}, captionpos=b,frame=single,numbers=none]
  $i_1$ = move r_t1 pc
@@
 
-The macro \texttt{mclear $r$} clears all the memory the
+The macro \texttt{mclear $r$} clears all the memory addresses that the
 capability in register $r$ has authority over. It is used by \texttt{scall} to
@@
 \texttt{prepstack} should be used when one receives a stack from an unknown
-source. The \texttt{prepstack} macro first ensures the stack has permission
-read/write-local/execute, and then it adjusts the stack capability, so it
+source. The \texttt{prepstack} macro first ensures the stack capability has permission
+read/write-local/execute. Then, it adjusts the stack capability, so it
 follows the convention for the stack\footnote{The stack capability should always
@@
 they help making the examples in \sectionname~\ref{sec:examples} intelligible.
-The macros \texttt{push $r$} and \texttt{pop $r$}
-add and remove elements from the stack. The macro \texttt{fetch
-    $r$ $\var{name}$} fetches the capability related to $\var{name}$ from the
-linking table and stores it in register $r$. The macro
-\texttt{malloc $r$ $n$} invokes malloc with size argument $n$.
-The \texttt{malloc} macro assumes that a capability for malloc resides in the
-linking table and is basically a fetch of the malloc capability followed by a
-setup of a return pointer. Finally, the macro \texttt{crtcls
+The macros \texttt{push $r$} and \texttt{pop $r$} respectively add and remove
+elements from the stack. The macro \texttt{fetch $r$ $\var{name}$} fetches the
+capability related to $\var{name}$ from the linking table and stores it in
+register $r$. The macro \texttt{malloc $r$ $n$} invokes malloc with size
+argument $n$. The \texttt{malloc} macro assumes that a capability for malloc
+resides in the linking table and is basically a fetch of the malloc capability
+followed by a setup of a return pointer. Finally, the macro \texttt{crtcls
   $\overline{(x_i,r_i)}$ $r$} allocates a closure where $r$ points to the
@@
 We stress that the macros correspond to series of instructions as seen in
-\figurename~\ref{fig:scall-impl}, and that they
+\figurename~\ref{fig:scall-impl}; the macros
 are introduced for intelligibility. The examples of
@@
 In that case, we now think it would suffice to clear only the (much smaller) part of the stack used by the trusted code itself.
-To understand this, it is useful to take another look at the illustrations in Figure~\ref{fig:ret-adv} related to this case.
+To understand this, it is useful to take another look at the illustrations in Figure~\ref{fig:ret-adv2} related to this case.
 If we do not clear the stack upon return (as illustrated in Figure~\ref{fig:illu-clear-all}), then the adversary might have used that stack to store local capabilities they received in a previous invocation (see Figure~\ref{fig:illu-adv-data}).

### 2019-02-10 toplas/paper.tex

@@
   \end{subfigure}
-\begin{subfigure}{0.043\linewidth}
+\begin{subfigure}{0.033\linewidth}
   % padding
@@
       \begin{scope}
-        \clip (4.6,5) rectangle (7.6,11.5);
-        \capbrace{(4.6,5)}{(4.6,13)}{sp}
+        \clip (4.6,7) rectangle (7.6,11.5);
+        \capbrace{(4.6,7)}{(4.6,13)}{sp}
       \end{scope}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.043\linewidth}
+\begin{subfigure}[t]{0.02\linewidth}
   % padding

### 2019-02-10 toplas/paper.tex
[TOPLAS] Added short paragraph about secure compilation

@@
 Building a logic around the semantic model presented here would remove some of the tedious repetitive proof details making it more realistic to prove properties about larger more realistic programs.
+
+If we want to scale this approach even further, we would like to reason at a
+higher level of abstraction namely at the level of a high-level language.
+That is, we would rather construct a program logic for a high-level language, so
+we can reason about the programs the language we actually write them in and at
+the same time get the guarantees for the programs we are going to execute, i.e.\
+the compiled programs.
+To achieve this, we would have to construct a secure compilation that preserves the security abstraction of the high-level language.
+Our logical relation would not be directly applicable in a secure compilation proof, the experiences made from constructing a logical relation for a capability machine can still be used.
 \end{toplas}
-\itoplassug{We could further add that this is not the level of abstraction that one would like to reason.
-One would like to reason about the programs we actually write - that is programs written in high-level languages.
-If we had a fully-abstract compiler from a nice high-level language to this low-level machine, then we could reason about the program in the high-level language and then because the compiler is fully-abstract, we would retain all the nice properties from the high-level language.
-How does this work fit in to this picture?
-The calling convention ensures two properties that we expect our nice high-level languages to have, namely well-bracketedness and local-state encapsulation.
-If we want to do fully-abstract compilation, then we need some way to guarantee this after compilation.
-Further, logical relations are often used in full-abstractness proofs.
-This work expands our knowledge about logical relations should be defined for low-languages and in particular capability machines.
-(maybe also mention that enforcement necessary for fully-abstract compilation.
-Could be capability, but it could also be something else - we just need something which we do not have now.)  }
+\itoplassug{We could further add that this is not the level of abstraction that
+  one would like to reason. One would like to reason about the programs we
+  actually write - that is programs written in high-level languages. If we had a
+  fully-abstract compiler from a nice high-level language to this low-level
+  machine, then we could reason about the program in the high-level language and
+  then because the compiler is fully-abstract, we would retain all the nice
+  properties from the high-level language. How does this work fit in to this
+  picture? The calling convention ensures two properties that we expect our nice
+  high-level languages to have, namely well-bracketedness and local-state
+  encapsulation. If we want to do fully-abstract compilation, then we need some
+  way to guarantee this after compilation. Further, logical relations are often
+  used in full-abstractness proofs. This work expands our knowledge about
+  logical relations should be defined for low-languages and in particular
+  capability machines. (maybe also mention that enforcement necessary for
+  fully-abstract compilation. Could be capability, but it could also be
+  something else - we just need something which we do not have now.) }
 

### 2019-02-10 toplas/paper.tex
[TOPLAS] Make small textual changes to introduction

@@
   pointers) and can be efficient assuming only one additional piece of processor
-  support: an efficient instruction for clearing a range of memory.
+  support (wrt CHERI): an efficient instruction for clearing a range of memory.
 \item We present a novel step-indexed Kripke logical relation for reasoning
@@
   This paper is an extension of the published conference paper~\citet{Skorstengaard:esop18}.
-  We have made improvements to readability and completeness throughout the paper, but we highlight the following changes:
+  We have made improvements to readability and completeness throughout the paper.
+For instance, we have added an introduction to \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
+We highlight the following changes:
   \begin{itemize}
-  \item An introduction to the \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
   \item Proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
@@
   \item A section on reasoning about programs that run on a capability machine.
-    This section explains how one reasons about common scenarios that arise in programs on a capability machine, and in particular how the logical relation is used.
+    This section explains how one reasons about common scenarios that arise in programs on a capability machine, and, in particular, how the logical relation is used.
     It also introduces a number of lemmas that prove recurring bits once and for all.
-  \item Details previously found only in the technical appendix~\citep{technical_appendix} to the paper.
-    Further, many of the still omitted details have been added to the paper's appendix.
+  \item Details previously found only in the technical appendix~\citep{technical_appendix} has been moved to the paper.
+    Further, many of the details omitted in the paper has been added to the paper's appendix.
   \end{itemize}

### 2019-02-10 toplas/paper.tex
[TOPLAS] Rewrite the syntactic/semantic discussion to not use those terms

@@
 \noindent \textbf{Reasoning about capability machine programs}\\
-\emph{Semantic, but not syntactic properties} The logical relation defined in
+\emph{Limitations} The logical relation defined in
 \sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
 programs. A limitation w.r.t. previous work is that the logical relation is
-tailored exclusively towards semantic properties, not syntactic ones.
+tailored towards a specific class of problems.
 
@@
 the code returns, we may try to prove that any capabilities passed back to us in
-the registers are still confined to that range of memory. The property of
-falling in a certain address range is syntactic in the sense that it talks about
-the specific implementation of a higher-order value rather than its behavior,
+the registers are still confined to that range of memory. The property talks
+about a the specific implementation of a higher-order value rather than its behavior,
 like the invariants that are required/preserved when we use it.
 
-Such syntactic properties are hard to prove in our system. For the example
-cited, it would be easy to conclude that the returned values are in the value
-relation (see \figurename~\ref{fig:logrel}). This gives us a lot of semantic
-information, like conditions under which they are safe to use and invariants
-that will be preserved when we do, but it does not tell us much about the
-address of the capability. As a very concrete example, capabilities with
-permission $\noperm$ are always in the value relation, irrespective of their
-address. Semantically, this makes perfect sense as they are always safe
-since they cannot be used for anything anyway. However, it also means we do not
-get syntactic information about them.
-
-For our purposes, this restriction is unproblematic since we are only
-interested in proving semantic properties (e.g., an assertion will never fail).
-However, in other situations, we may be interested in proving more syntactic
-properties like the ones that are often considered in object capability
-literature: confinement, no authority amplification etc. Although such
-properties are more restrictive and less easy to use for reasoning,
-\citet{Devriese:2016ObjCap} have demonstrated how a logical relation like
-ours can be adapted to also support them, by quantifying the logical relation
-over a custom interpretation of effectful computations and the type of
-references. We expect their solution can be readily adapted to our setting,
-modulo some details (like the fact that we do not just have read-write
-capabilities, but also others).
+Such properties are hard to prove in our system.
+For the example, it would be easy to conclude that the returned values are in the value relation (see \figurename~\ref{fig:logrel}).
+This gives us a lot of behavioral information, like conditions under which they are safe to use and invariants that will be preserved when we do, but it does not tell us much about the address of the capability.
+As a very concrete example, capabilities with permission $\noperm$ are always in the value relation irrespective of their address.
+Behaviorally, this makes perfect sense; $\noperm$-capabilities cannot be used
+for anything.
+
+For our purposes, this restriction is unproblematic since we are only interested in proving behavioral properties (e.g., an assertion will never fail).
+In other situations, however, we may be interested in proving properties like the ones that are often considered in object capability literature: confinement, no authority amplification etc.
+Although such properties are more restrictive and less easy to use for reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation like ours can be adapted to also support them by quantifying the logical relation over a custom interpretation of effectful computations and the type of references.
+We expect their solution can be readily adapted to our setting modulo some details (like the fact that we do not just have read-write capabilities but also others).
 \end{toplas}

### 2019-02-10 toplas/paper.tex
[TOPLAS] Add secure compilation reference

@@
 the same time get the guarantees for the programs we are going to execute, i.e.\
-the compiled programs.
-To achieve this, we would have to construct a secure compilation that preserves the security abstraction of the high-level language.
+the compiled programs. To achieve this, we would have to construct a secure
+compilation \citep{patrignani_secure_compilation_2019} that preserves the security abstraction of the high-level language.
 Our logical relation would not be directly applicable in a secure compilation proof, the experiences made from constructing a logical relation for a capability machine can still be used.

### 2019-02-10 toplas/paper.tex
[TOPLAS] Add discussion on lifting to realistic machine

@@
 It is inspired by both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and CHERI~\citep{Watson2015Cheri}.
-To avoid tedious details irrelevant to the focus of this paper, we assume an infinite address space and unbounded integers.
+To avoid tedious details irrelevant to the focus of this paper, we assume an
+infinite address space and unbounded integers (see
+\sectionname~\ref{sec:discussion} for more on this).
 
@@
 object-oriented program or every closure in a functional program.
+\\\\
+\begin{toplas}
+  \noindent \textbf{Capability machine formalization}\\
+  \emph{Simplifications} Our capability machine formalization assumes unbounded integers and an infinite address space.
+  Further, it has a much simpler instruction set than that of a full ISA.
+  By making these simplifications, we avoid tedious details but end up with a less realistic machine.
+  However, the intend of this work is to gain ground in the formal work necessary to prove properties about low-level assembly programs on a capability machine not to apply it to a full fledged capability machine.
+
+  We do not believe that any of our simplifications are beyond reason.
+  If we added unbounded integers, we would have to change the operational
+  semantics to take over- and under flows into account. This could be achieved
+  by changing the \texttt{plus} and \texttt{minus} instruction to use modulo
+  arithmetic.
+  If we added a bounded address space, then we would have to take memory out of
+  bounds into account. This would require changes to the memory
+  operations (\texttt{store} and \texttt{load}) and the step relation.
+  The malloc specification would also need to be updated as it would have to
+  fail when it runs out of memory.
+  Finally, expanding the instruction set to a realistic ISA would add a handful
+  of instructions that we would need to reason about.
+  All in all, it would add a lot of tedious work to expand this result to a
+  realistic machine. The amount of details in the proofs are already at the
+  threshold for what should be done with pen and paper, so expanding this work
+  to a realistic machine would require mechanized proofs that can take care of
+  the tedious details.
+\end{toplas}
 \\\\

### 2019-02-10 toplas/paper.tex
[TOPLAS] Add reference to attack-defense arms race
... also fix reference.

@@
 Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
-As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-mitigation arms race) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
+As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-defense arms race \citep{Szekeres_sok_2013}) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
 The latter techniques essentially emulate protection techniques on existing hardware at the cost of performance, system complexity, and/or security.
 
-\emph{Capability machines} are a type of processors that
-remediate these limitations with a better security model at
-the hardware level. They are based on old
-ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that
- have recently received renewed interest. in particular, the CHERI
-project has proposed new ideas and ways of tackling practical
-challenges like backwards compatibility and realistic OS
-support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}. Capability
-machines tag every word (in the register file and in memory) to
-enforce a strict separation between numbers and capabilities (a kind
-of pointers that carry authority). Memory capabilities carry
-the authority to read and/or write to a range of memory
-locations. There is also a form of \emph{object capabilities}, which represent the
-authority to invoke a piece of code without exposing the code's
-encapsulated private state (e.g., the M-Machine's enter capabilities or
-CHERI's sealed code/data pairs).
+\emph{Capability machines} are a type of processors that remediate these limitations with a better security model at the hardware level.
+They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that have recently received renewed interest.
+in particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
+Capability machines tag every word (in the register file and in memory) to enforce a strict separation between numbers and capabilities (a kind of pointers that carry authority).
+Memory capabilities carry the authority to read and/or write to a range of memory locations.
+There is also a form of \emph{object capabilities}, which represent the authority to invoke a piece of code without exposing the code's encapsulated private state (e.g., the M-Machine's enter capabilities (described in Section~\ref{sec:capab-mach-with}) or CHERI's sealed code/data pairs).
 
@@
 (ideally evidenced by a mathematical proof, guaranteeing that we do not end up
-in a new attack-mitigation arms race). However, a lot needs to happen before we get
+in a new attack-defense arms race). However, a lot needs to happen before we get
 there. For example, it is far from trivial to devise a compilation scheme

### 2019-02-10 toplas/paper.tex
[TOPLAS] Added a short reference to our POPL paper

@@
 
-If we want to scale this approach even further, we would like to reason at a
-higher level of abstraction namely at the level of a high-level language.
-That is, we would rather construct a program logic for a high-level language, so
-we can reason about the programs the language we actually write them in and at
-the same time get the guarantees for the programs we are going to execute, i.e.\
-the compiled programs. To achieve this, we would have to construct a secure
-compilation \citep{patrignani_secure_compilation_2019} that preserves the security abstraction of the high-level language.
-Our logical relation would not be directly applicable in a secure compilation proof, the experiences made from constructing a logical relation for a capability machine can still be used.
+If we want to scale this approach even further, we would like to reason at a higher level of abstraction namely at the level of a high-level language.
+That is, we would rather construct a program logic for a high-level language, so we can reason about the programs the language we actually write them in and at the same time get the guarantees for the programs we are going to execute, i.e.\ the compiled programs.
+To achieve this, we would have to construct a secure compilation \citep{patrignani_secure_compilation_2019} that preserves the security abstraction of the high-level language.
+The experiences made about logical relations for capability machines here has been used by \citet{skorstengaard_stktokens_2019} to prove make advances in the direction of building a secure compilation targeting capability machines.
+They present a calling convention based on linear capabilities rather than local capabilities, and they provide a fully abstract compilation that can be used as the last phase in a secure compilation.
+The full-abstraction proof uses a logical relation.
 \end{toplas}

### 2019-02-10 toplas/paper.tex
[TOPLAS] Add notes, update TODO and fix reference
[TOPLAS] Add CHERI reference for stack smashing

@@
 stack overwriting activation records' contents). Luckily, the capability
-machine's fine-grained memory protection should make it reasonably easy for a
+machine's fine-grained memory protection \citep{Woodruff:2014:CCM:2665671.2665740} should make it reasonably easy for a
 compiler to implement such protection by making sure that only appropriately
@@
 specific programs that rely on well-bracketed control flow.
+
+% \citet{skorstengaard_stktokens_2019} gives a definition of well-bracketed control flow based on overlay semantics.
+% Their definition characterizes the kind of well-bracketed control flow our examples rely on.
+% They define well-bracketed control flow for a capability machine by defining a new machine that is syntactically the same but adds a built-in call stack that is only accessed when a series of instructions corresponding to a call or return is executed.
 

### 2019-02-11 toplas/paper.tex
[TOPLAS] Update reasoning about cap machine section

@@
   pointers) and can be efficient assuming only one additional piece of processor
-  support (wrt CHERI): an efficient instruction for clearing a range of memory.
+  support (w.r.t. CHERI): an efficient instruction for clearing a range of memory.
 \item We present a novel step-indexed Kripke logical relation for reasoning
@@
 % }
-There are many details to get right when programming in assembly. These details
-carry over to proofs about assembly programs, so many of the proofs in the next
-section about example programs are a bit cumbersome. It is especially annoying
-when the same line of reasoning is applied in multiple places. To mitigate this,
-we have defined a number of lemmas that capture common reasoning patterns in
-these proofs. We present the most important such lemmas in this section, to give
-the reader an idea of how they work.
+There are many details to get right when programming in assembly.
+These details carry over to proofs about assembly programs, so many of the proofs in \sectionname~\ref{sec:examples} about example programs are a bit cumbersome.
+It is especially annoying when the same line of reasoning is applied in multiple places.
+To mitigate this, we have define a number of lemmas that capture common reasoning patterns in these proofs.
+In this section, we present the most central lemmas used to reason about assembly programs.
 
@@
 Our capability machine only allows the final memory of an execution to be observed, so naturally the correctness lemmas we prove in \sectionname~\ref{sec:examples} are statements about the memory in the halted configuration.
-In order to prove something about some part of the final memory, we establish an invariant for it as a region in a world and show that the initial configuration is in the $\observations$-relation w.r.t.\ that world.
-The invariants we are interested in must be permanent regions so that they are not revoked during execution.
+In order to prove a property about part of the final memory, we create a world with a region the ensures the desired property and show that the initial configuration is in the $\observations$-relation w.r.t.\ that world.
+The region that ensures the property must be permanent, so it is not revoked during execution.
 The $\observations$-relation says that if the initial configuration halts successfully, then the memory in the halting state must still respect a private future world of the initial world.
 It is, however, bothersome and error prone to try to argue about the entire execution in one go.
-Instead we want to reason modularly, in the sense that we only want to reason about parts of the execution at a time.
-To this end, we prove an anti-reduction lemma that essentially says that if we can show that an initial configuration, $\Phi$, steps to a configuration $\Phi'$ and that $\Phi'$ is in the $\observations$-relation, then $\Phi$ is also in the observation relation.
+Instead we want to reason modularly in the sense that we only want to reason about parts of the execution at a time.
+To this end, we prove an anti-reduction lemma that essentially says that if we can show that an initial configuration $\Phi$ steps to a configuration $\Phi'$ and $\Phi'$ is in the $\observations$-relation, then $\Phi$ must also be in the observation relation.
 \begin{lemma}[Anti-reduction for $\observations$]
@@
 % FTLR
-The anti-reduction works when we reason about the execution of known code,
-because we can execute it. When we want to reason about unknown code, the
-anti-reduction lemma does not apply because we do not know what instructions are
-being executed. This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
-As a reminder, the FTLR says that if a capability only has access to safe values
-with respect to a world, then it is safe to use it for execution in the same
-world. The unknown code we consider will be assumed to only have access to safe
-values, which typically means that we assume it has access to a linking table
-with a malloc capability and otherwise consists of instructions. This allows us
-to use the FTLR seeing as malloc is a safe value (cf.
-Lemma~\ref{lem:malloc-in-vr}) and instructions are really just integers and they
-are always safe values. Here, it is important to remember that the FTLR gives us
-that the capability for the untrusted code is safe w.r.t. some world, so even
-though we described the usual assumption on the unknown code's memory, it is the
-semantic model of the memory, i.e.\ the world, that the FTLR considers. In order
-to argue that a specific configuration is safe, we need to argue that the
-configuration is safe with respect to the world. That is, we need to argue that
-the memory satisfies the world and the register-file is in the
-$\stdrr$-relation. For instance, in the case where the untrusted code assumes
-control first, we will use the FTLR to show that the capability for the unknown
-code can be used for execution. The unknown code will get access to some known,
-trusted code through a capability in the initial register-file. Because the
-capability is in the register-file, we will have to argue that it is safe which
-we do using Lemma~\ref{lem:anti-red-obs}.
+We use the anti-reduction lemma when we reason about the execution of known code.
+When we want to reason about unknown code, the anti-reduction lemma does not apply because we do not know what instructions are being executed.
+This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
+As a reminder, the FTLR says that if a capability only has access to safe values with respect to a world, then it is safe to use the capability for execution in the same world.
+The unknown code we consider will be assumed to only have access to safe values which typically means it has access to a linking table with a malloc capability and otherwise consists of instructions.
+These assumptions allow us to use the FTLR to reason about the unknown code as malloc is a safe value (cf.\ Lemma~\ref{lem:malloc-in-vr}) and instructions are integers; integers are always safe values.
+It is important to remember that safety is relatively to the semantic model of a memory and not the actual memory itself.
+For this reason, the assumptions we make on unknown code must be expressed as a region in a world.
+
+In order to argue that a specific configuration is safe, we need to argue that the configuration is safe with respect to the world.
+That is, we need to argue that the memory satisfies the world and the register-file is in the $\stdrr$-relation.
+For instance in the case where the untrusted code assumes control first, we will use the FTLR to show that the capability for the unknown code can be used for execution.
+The unknown code will get access to some known, trusted code through a capability in the initial register-file, so we will have to argue that the known code is safe.
+We argue about the known code with Lemma~\ref{lem:anti-red-obs}.
 
 % Many details to get right when reasoning about low-level programs.
-Another common pattern in proofs about programs on our capability machine
-concerns the use of scall. The following lemma captures the commonalities of
-reasoning about programs using scall. From another point of view, it can be seen
-as a specification for scall.
+Another common pattern in proofs about programs on our capability machine concerns the use of \texttt{scall}.
+The following lemma captures the commonalities of reasoning about programs using \texttt{scall}.
+For instance, setting up the local stack frame and constructing stack pointers and protected return pointers for the callee always amount to the same line of reasoning which is captured by the lemma.
+From another point of view, it can be seen as a specification for \texttt{scall}.
 \begin{lemma}[$\mathtt{scall}$ works]
@@
 
-    then we have that $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
+    then we have $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
   \item \label{item:scall:hyp-cont} \emph{Hyp-Cont} For all $n' \in \nats$, $W''
@@
     \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
-    \item for all $r$, we have that:
+    \item for all $r$, we have:
       \begin{equation*}
@@
 
-    then we have that $\npair[n']{(\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')} \in \observations(W'')$
+    then we have $\npair[n']{(\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')} \in \observations(W'')$
   \end{enumerate}
@@
 \end{lemma}
-\begin{figure}[htb]
+\begin{toplas}
+  The \texttt{scall} lemma is stated in terms of a number of auxiliary definitions found in the appendix.
+  The region $\iota^{sta}(\gl,\ms)$ is a static region with locality $\gl$.
+  It is static in the sense that it only accepts the memory segment $\ms$ (\appendixname~\ref{app:awkward-example}).
+  The function $\revokeTemp{} : \Worlds \fun \Worlds$ yields the input world but with all the temporary regions replaced with revoked regions (\appendixname~\ref{app:logical-relation}).
+  The definition of \textit{$(\reg,\ms)$ is looking at $[i_0\dots i_n]$ followed by $c_{\var{next}}$} is visualized in \figurename~\ref{fig:def-intuition-looking-at}; it requires the capability in $\pcreg$ of $\reg$ to point to the first address of the instructions $[i_0\dots i_n]$ in $\ms$ and $c_{\var{next}}$ to point to the address immediately after the instructions.
+  The definition of \textit{$\reg$ points to stack with $\ms_\stk$ used and $\ms_{\stk,\var{unused}}$ unused} is visualized in \figurename~\ref{fig:def-intuition-points-to-stack}; it requires $\ms_\stk$ and $\ms_{\stk,\var{unused}}$ to be adjacent and continuous memory segments with the local $\rwlx$-capability in \texttt{r\_stk} of $\reg$ governing the two memory segments and pointing to the top most address of $\ms_\stk$ (\appendixname~\ref{app:programs-definitions}).
+\end{toplas}
+
+Roughly, the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}) states that an invocation of scall is safe if \texttt{scall} is executed in a reasonable state (\ref{item:scall:memsat}-\ref{item:scall:stack}), the callee is safe to execute (\ref{item:scall:hyp-callee}), and returning to the code after the scall is safe (\ref{item:scall:hyp-cont}).
+
+In the common case, \texttt{scall} is used to reason about untrusted code that we only have general assumptions about.
+As explained previously, we use the FTLR (Theorem~\ref{thm:ftlr}) to argue about unknown code; this is no exception.
+That is, when we argue that we assumption~\ref{item:scall:hyp-callee} in the \texttt{scall} lemma satisfied, we use the FTLR along with the assumptions we have (e.g.\ linking table, malloc capability etc.).
+In Hyp-Callee, we assume the memory is safe, so it suffices to show that the register-file contains safe values, which amounts to showing that the arguments in the call are safe.
+
+By using Lemma~\ref{lem:scall-works} to reason about \texttt{scall}, the proofs become agnostic to the actual implementation of \texttt{scall}.
+In other words, should we change the implementation of \texttt{scall}, then we just need to prove that Lemma~\ref{lem:scall-works} holds for the new implementation in order to get that all our results still hold true.
+
+\begin{figure}[bt]
   \centering
@@
     \caption{Visualization of ``$(\reg,\ms)$ is looking at $[i_0\dots i_n]$
-      followed by $c_{\var{next}}$''.}
+      followed by $c_{\var{next}}$''. $\reg(\pcreg)$ points to $i_0$ and
+      $c_{\var{next}}$ points to the address after $i_n$}
   \label{fig:def-intuition-looking-at}
@@
     \caption{Visualization of ``$\reg$ points to stack with $\ms_\stk$ used and
-      $\ms_{\stk,\var{unused}}$ unused''.}
+      $\ms_{\stk,\var{unused}}$ unused''. $\reg(r_{\var{stk}})$ is a local
+      $\rwlx$-capability that points to the
+      top most address of $\ms_\stk$.}
   \label{fig:def-intuition-points-to-stack}
@@
 
-  \caption{
-  }
+  \caption{}
   \label{fig:def-intuition}
 \end{figure}
-\begin{toplas}
-The \texttt{scall} lemma is stated in terms of a number of auxiliary definitions
-found in the appendix. The region $\iota^{sta}(\gl,\ms)$ is a static region with
-locality $\gl$. It is static in the sense that it only accepts the region $\ms$
-(\appendixname~\ref{app:awkward-example}).
-The function $\revokeTemp{} : \Worlds \fun \Worlds$ yields the input world but
-with all the temporary regions replaced with revoked regions (\appendixname~\ref{app:logical-relation}). The definition of \textit{$(\reg,\ms)$ is looking at $[i_0\dots i_n]$
-      followed by $c_{\var{next}}$} is visualized in
-    \figurename~\ref{fig:def-intuition-looking-at};
-    it requires the capability in $\pcreg$ of
-    $\reg$ to point to the first address of the instructions $[i_0\dots i_n]$ in
-    $\ms$ and $c_{\var{next}}$ to point to the address immediately after the
-    instructions. The definition of \textit{$\reg$ points to stack with $\ms_\stk$ used and
-      $\ms_{\stk,\var{unused}}$ unused} is visualized in
-    \figurename~\ref{fig:def-intuition-points-to-stack};
-    it requires $\ms_\stk$ and $\ms_{\stk,\var{unused}}$ to be adjacent and
-    continuous memory segments with the local $\rwlx$-capability in
-    \texttt{r\_stk} of $\reg$ governing the two memory segments and pointing to
-    the top most address of $\ms_\stk$ (\appendixname~\ref{app:programs-definitions}).
-\end{toplas}
-
-Roughly, the \texttt{scall} lemma states that an invocation of scall is safe if
-\texttt{scall} is executed in a reasonable state
-(\ref{item:scall:memsat}-\ref{item:scall:stack}), the callee is safe to execute
-(\ref{item:scall:hyp-callee}), and returning to the code after the scall is safe
-(\ref{item:scall:hyp-cont}).
-
-In the common case, \texttt{scall} is used to invoke untrusted code, and we will
-only have very general assumptions about that code, in which case we will use
-the FTLR (Theorem~\ref{thm:ftlr}) to argue
-Assumption~\ref{item:scall:hyp-callee}, using what we know about the values it
-has access to (e.g.\ linking table, malloc capability etc.). By assumption in
-Hyp-Callee, the memory is safe, so it suffices to show that the register-file
-contains safe values, which amounts to showing that the arguments in the call
-are safe.
-
-By using Lemma~\ref{lem:scall-works} to reason about \texttt{scall}, the proofs
-become agnostic to the actual implementation of \texttt{scall}. In other words,
-should we change the implementation of \texttt{scall}, then we just need to
-prove that Lemma~\ref{lem:scall-works} holds for the new implementation in order
-to gain that all our results still hold true.
-
-Also often used is the \texttt{malloc} instruction, so we also prove
-Lemma~\ref{lem:malloc-works} to help reason about it. The structure of the
-\texttt{malloc} lemma is close to that of the \texttt{scall} lemma. It also has
-requirements on the configuration just before \texttt{malloc} is executed
-(\ref{item:malloc:looking-at}-\ref{item:malloc:malloc-sat}) as well as
-requirements on the execution afterwards (\ref{item:malloc:hyp-cont}). It does
-not have any requirements on the callee as the \texttt{malloc} specification
-defines its behavior.
+
+The \texttt{malloc} macro is also common in our examples, so we prove Lemma~\ref{lem:malloc-works} to help reason about it.
+The structure of the \texttt{malloc} lemma is close to that of the \texttt{scall} lemma.
+It also has requirements on the configuration just before \texttt{malloc} is executed (\ref{item:malloc:looking-at}-\ref{item:malloc:malloc-sat}) as well as requirements on the execution afterwards (\ref{item:malloc:hyp-cont}).
+It does not have any requirements on the callee as the \texttt{malloc} specification defines its behavior.
 \begin{lemma}[$\mathtt{malloc}$ works]
@@
 \end{lemma}
-In the technical appendix~\citep{technical_appendix}, we also define a lemma for
-the macro used to create closures, \texttt{crtcls}.
+In the technical appendix~\citep{technical_appendix}, we also define a lemma for the macro used to create closures, \texttt{crtcls}.
+
 % The below remark seems unnecessary: isn't this true about any lemma in any proof?
+
 % We stress that these lemmas
 % are conveniences and are not strictly necessary to reason about programs for the
-% capability machine. In principle, one could reason solely in the semantic model
+% capability machine.
+In principle, one could reason solely in the semantic model
 % without the use of these lemmas, but that would lead to a lot of repetition.

### 2019-02-12 toplas/paper.tex
[TOPLAS] Make small textual changes to examples section

@@
   move r_t1 pc
-  lea r_t1 $\var{ofc}$ 
+  lea r_t1 $\var{off}$ 
   push r_t1
@@
   move r_0 r_stk
-  lea r_0 $\var{ofc_\var{rec}}$
+  lea r_0 $\var{off_\var{rec}}$
   restrict r_0 $\encodePermPair((\local,\entry))$
@@
  $i_1$ = move r_t1 pc
- $i_2$ = lea r_t1 $\var{ofc}_\var{stk}$
+ $i_2$ = lea r_t1 $\var{off}_\var{stk}$
  $i_3$ = load r_stk r_t1
@@
       restoration code is presented in the top right corner. The
-      variable $\var{ofc}_\var{ret}$ is the offset to the label \texttt{return}, and
-      $\var{ofc_\var{rec}} = -5$ which is the offset to the first instruction of
+      variable $\var{off}_\var{ret}$ is the offset to the label \texttt{return}, and
+      $\var{off_\var{rec}} = -5$ which is the offset to the first instruction of
       the activation record. The set $R = \RegName \setminus
       \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.
-    The variable $\var{ofc}_\var{stk} =5$ which is the offset to the address
+    The variable $\var{off}_\var{stk} =5$ which is the offset to the address
     where the old stack pointer is stored on the stack.}
@@
 \itoplassug{Suggestion: add a couple of short proof sketches.}
-In this section, we demonstrate how our formalization of capability safety
-allows us to prove local-state encapsulation and control-flow correctness
-properties for challenging program examples. The security measures
-of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to ensure these
-properties. Since we are dealing with assembly language, there are many details
-to the formal treatment, and therefore we necessarily omit some details in the
-lemma statements. The examples may look deceivingly short, but it is because
-they use the macro instructions described in \sectionname~\ref{sec:macros}.
-The examples would be
-unintelligible without the macros, as each macro expands to multiple basic instructions. The interested reader can find all the
-technical details in the technical appendix~\citep{technical_appendix}.
+In this section, we demonstrate how our formalization of capability safety allows us to prove local-state encapsulation and control-flow correctness properties for challenging program examples.
+The security measures of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to ensure these properties.
+Since we are dealing with assembly language, there are many details to the formal treatment. Therefore, we omit some details in the lemma statements.
+The examples may look deceivingly short, but it is because we use the macro instructions described in \sectionname~\ref{sec:macros}.
+The examples would be unintelligible without the macros as each macro expands to multiple basic instructions.
 
 \subsection{Encapsulation of Local State}
-\texttt{f1} and \texttt{f2} in The programs in \figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's encapsulation of local state.
-They are very similar: both store some local state, call an untrusted piece of code ($\adv$), and then test whether the local state is unchanged.
+The programs \texttt{f1} and \texttt{f2} in \figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's encapsulation of local state.
+They are very similar: both store some local state, call an untrusted piece of code ($\adv$), and test whether the local state is unchanged after the call.
 They differ in the way they do this.
@@
 
-For both programs, we can prove that if they are linked with an adversary,
-$\adv$, that is allowed to allocate memory but has no other capabilities, then
-the assertion will never fail during executing (see
+For both programs, we prove that if they are linked with an adversary,
+$\adv$, allowed to allocate memory but has no other capabilities, then
+the assertion will never fail during execution (see
 Lemmas~\ref{lem:correctness-f1} and~\ref{lem:correctness-f2} below).
@@
 The logical relation is not specific to any calling convention, so we can use it to reason
-about both programs, even though they use different calling conventions.
+about both programs even though they use different calling conventions.
 
@@
 fails. To this end, we assume they have access to a flag (an address in memory).
-If the assertion fails, then the flag is set to $1$ and execution halts. The correctness lemma
-for \texttt{f1} then states:
-\begin{lemma}
+If the assertion fails, then the flag is set to $1$ and execution halts. 
+\begin{lemma}[\texttt{f1} is correct]
   \label{lem:correctness-f1}
@@
 
-To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
-configuration is safe (in the $\observations$ relation) for a world with a
-permanent region that requires the assertion flag to be 0. By
-Lemma~\ref{lem:anti-red-obs}, it suffices to show that the configuration is safe
-after some reduction steps. We then use the \texttt{scall} lemma
-(Lemma~\ref{lem:scall-works}), by which it suffices to show that (1) the
-configuration that \texttt{scall} will jump to is safe and (2)
-that the configuration just after \texttt{scall} is done cleaning
-up is safe. We use the Fundamental Theorem to reason about the unknown
-adversarial code as described in \sectionname~\ref{sec:reasoning}, but notice
-that the adversary capability is an enter capability, which the Fundamental
-Theorem says nothing about. Luckily the enter capability becomes $\exec$ after
-the jump and then the Fundamental Theorem applies.
+To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start configuration is safe (in the $\observations$ relation) for a world with a permanent region that requires the assertion flag to be 0.
+By Lemma~\ref{lem:anti-red-obs}, it suffices to show that the configuration is safe after some reduction steps.
+We then use the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}) by which it suffices to show that (1) the configuration that \texttt{scall} will jump to is safe and (2) the configuration just after \texttt{scall} is done cleaning up is safe.
+We use the Fundamental Theorem to reason about the unknown adversarial code as described in \sectionname~\ref{sec:reasoning}, but notice that the adversary capability is an enter capability, which the Fundamental Theorem says nothing about.
+Luckily, the enter capability has $\exec$-permission after the jump at which point the Fundamental Theorem applies.
 
 We have a similar lemma for \texttt{f2}:
-\begin{lemma}
+\begin{lemma}[\texttt{f2} is correct]
   \label{lem:correctness-f2}
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer,
+  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer and assuming $c_{f2}$ points to \texttt{f2},
   if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
@@
 \subsection{Well-Bracketed Control-Flow} 
-Using the stack-based calling convention of \texttt{scall}, we get
-well-bracketed control-flow. To illustrate this, we look at
-two example programs \texttt{f3} and
-\texttt{g1} in \figurename~\ref{fig:prog-f3-and-g1}.
-
-In \texttt{f3} there are two calls to an adversary and in order
-for the assertion in the middle to succeed, they need to be well-bracketed. If
-the adversary were able to store the return pointer from the first call and
-invoke it in the second call, then \texttt{f3} would have $2$ on
-top of its stack and the assertion would fail. However, the security measures in
-\sectionname~\ref{sec:stack-and-return-pointer} prevent this attack: specifically,
-the return pointer is local, so it can only be stored on the stack, but the part
-of the stack that is accessible to the adversary is cleared before the second
-invocation. In fact, the following lemma shows that there are also no other
-attacks that can break well-bracketedness of this example, i.e.\ the assertion
-never fails. It is similar to the two previous lemmas:
+The stack-based calling convention \texttt{scall} ensures well-bracketed control-flow.
+This is illustrated by program examples \texttt{f3} and \texttt{g1} in \figurename~\ref{fig:prog-f3-and-g1}.
+
+The program \texttt{f3} has two calls to an adversary.
+In order for the assertion on line 5 to succeed, the calls must be well-bracketed.
+If the adversary were able to store the return pointer from the first call and invoke it in the second call, then \texttt{f3} would have $2$ on top of its stack and the assertion would fail.
+However, the security measures in \sectionname~\ref{sec:stack-and-return-pointer} prevent this attack.
+Specifically, the return pointer is local, so it can only be stored on the stack.
+However, the part of the stack that is accessible to the adversary is cleared before the second invocation preventing attempts to store the return pointer.
+In fact, the following lemma shows that there are also no other attacks that can break well-bracketedness of this example, i.e.\ the assertion never fails.
+It is similar to the two previous lemmas:
 
@@
      move $r_3$ $\pcreg$
-     lea $r_3$ $\var{offset}$
+     lea $r_3$ $\var{off}$
      crtcls $[(x, r_2)]$ $r_3$
@@
   \centering
+  \begin{minipage}[t]{.25\linewidth}
+  \begin{lstlisting}
+f3: push 1
+    fetch $r_1$ $\adv$
+    scall $r_1$($[]$,$[r_1]$)
+    pop $r_2$
+    assert $r_2$ 1
+    push 2
+    scall $r_1$($[]$,$[]$)
+    halt
+\end{lstlisting}
+  \end{minipage}%
   \begin{minipage}[t]{.37\linewidth}
@@
      move $\pcreg$ $r_3$
-     lea $r_3$ $\var{offset}$
+     lea $r_3$ $\var{off}$
      crtcls $[(x, r_2)]$ $r_3$
@@
 \end{lstlisting}
-  \end{minipage}%
-  \begin{minipage}[t]{.25\linewidth}
-  \begin{lstlisting}[firstnumber=last]
-f3: push 1
-    fetch $r_1$ $\adv$
-    scall $r_1$($[]$,$[r_1]$)
-    pop $r_2$
-    assert $r_2$ 1
-    push 2
-    scall $r_1$($[]$,$[]$)
-    halt
-\end{lstlisting}
   \end{minipage}
   \caption{ Two programs that rely on well-bracketedness of
-    \texttt{scall}s to function correctly. $\var{offset}$ is the
+    \texttt{scall}s to function correctly. $\var{off}$ is the
     offset to \texttt{f4}.}
@@
   Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1}
+  Lemma~\ref{lem:correctness-f1} and assuming $c_{f3}$ points to \texttt{f3}
   if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
@@
 
-The final example, \texttt{g1} with \texttt{f4},
-is a faithful translation of a tricky example known from the literature (known
-as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists
-of two parts, \texttt{g1} and \texttt{f4}.
-\texttt{g1} is a closure generator that generates closures with
-one variable $x$ set to $0$ in its environment and \texttt{f4} as
-the program (note we can omit some calling convention security measures because
-the stack is not used in the closure generator). \texttt{f4}
-expects one argument, a callback. It sets $x$ to $0$ and calls the callback.
-When it returns, it sets $x$ to $1$ and calls the callback a second time. When
-it returns again, it asserts $x$ is $1$ and returns. This example is more
-complicated than the previous ones because it involves a closure invoked by the
-adversary and an adversary callback invoked by us.
-\itoplassug{
-  Suggestion: Maybe give some more intuition as to why it is difficult.
-}
-As explained in
-\sectionname~\ref{sec:stack-and-return-pointer}, this means that we need to check (1)
-that the stack pointer that the closure receives from the adversary has
-write-local permission and (2) that the adversary callback is global.
+The final example, \texttt{g1} with \texttt{f4}, is a faithful translation of a tricky example known from the literature (known as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}.
+It consists of two parts, \texttt{g1} and \texttt{f4}.
+The program \texttt{g1} is a closure generator that generates closures with one variable $x$ set to $0$ in its environment and \texttt{f4} as the program (note we omit some calling convention security measures because the stack is not used in the closure generator).
+The program\texttt{f4} expects one argument, a callback.
+\texttt{f4} sets $x$ to $0$ and calls the callback.
+When the callback returns, \texttt{f4} sets $x$ to $1$ and calls the callback again.
+When the callback returns again, \texttt{f4} asserts $x$ is $1$ and returns.
+This example is more complicated than the previous ones because it involves a closure invoked by the adversary and an adversary callback invoked by us.
+\itoplassug{ Suggestion: Maybe give some more intuition as to why it is difficult.
+} As explained in \sectionname~\ref{sec:stack-and-return-pointer}, we need to check that (1) the stack pointer that the closure receives from the adversary has write-local permission and (2) the adversary callback is global.
 
@@
 store a stack pointer or trick the trusted code into preserving it across an
-invocation, the assertion can likely be made to fail too. However, our calling
+invocation, the assertion could be made to fail too. However, our calling
 convention prevents any of this from happening, as we prove in the following
@@
     \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
-    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\stk}$
+    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\adv}$
     \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
@@
   In order to get \eqref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
-  We have defined the world, so there (almost) is a one-to-one correspondence between region and memory segment in the initial configuration, so (a) is easily shown.
+  We have defined the world, so there (almost) is a one-to-one correspondence between regions in the world and memory segments in the initial configuration, so (a) easily follows.
   In order to show (b), we need to show $c_\stk \in \stdvr(W)$ and $c_{g1} \in \stdvr(W)$.
@@
 
-  The next part in the execution is an \texttt{scall}, so according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
+  The next part in the execution is an \texttt{scall}, so, according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
   For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
@@
   In $W_2$, two new regions govern the stack.
-  One of the new two regions governs the caller stack frame, i.e.\ the part of the stack that contains the contents of private registers and stack recovery code.
+  One of the two new regions governs the caller stack frame, i.e.\ the part of the stack that contains the contents of private registers and stack recovery code.
   The region that takes care of this is a temporary static region which ensures that our local stack frame remains unchanged during the execution of the callback.
-  The other new region we add takes care of the unused part of the stack which we are going to let the callback use for its execution.
+  The other new region takes care of the unused part of the stack which we are going to let the callback use for its execution.
   This is taken care of by a standard region $\iota^{pwl}$-region which allows the callback to store any safe value on the stack.
-  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we may assume that the arguments for the invocation of $c_{f4}$ are safe).
+  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we assume that the arguments for the invocation of $c_{f4}$ are safe).
   Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
@@
   This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still present in $W_3$.
-  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 1), so for the part of the execution up until the \texttt{scall} we once again use the anti-reduction lemma.
-
-  For the \texttt{scall} itself we, of course, apply the \texttt{scall} lemma.
+  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 1).
+  We use the anti-reduction lemma to reason about the execution until the second \texttt{scall}.
+
+  For the \texttt{scall} itself we apply the \texttt{scall} lemma.
   As the private stack has changed, we need to replace the two regions that govern the stack.
   This means that the second invocation of the callback takes place in a world $W_4 \futurestr W_3$.
-  The reasoning about \texttt{scall} is the same as for the first callback invocation.
+  The reasoning about \texttt{scall} is similar to the first callback invocation.
   The callback capability is still safe because it is global which basically covers the Hyp-Callee case.
@@
   As we used the anti-reduction lemma to get to this point, $W_6$ must also be constructed to be a private future world of $W_5$.
-  Before we construct $W_6$, let us consider where each of the worlds we have seen so far came from: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we constructed $W_2$ and $W_4$ which are the only private future worlds we have considered so far.
+  Before we construct $W_6$, let us consider where each of the worlds we have seen so far came from: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we constructed $W_2$ and $W_4$.
+  These are the only private future worlds we have considered so far.
   This means that we know exactly what changes they made.
@@
   In $W_5$ there were regions to handle the stack.
-  These regions obviously need to be revoked as the $W_1$ stack regions replace them.
+  These regions need to be revoked as the $W_1$ stack regions replace them.
   The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world relation permits as it is extensional.
@@
   move r_t2 pc
-  lea r_t2 $\var{ofc}_\var{end}$ 
+  lea r_t2 $\var{off}_\var{end}$ 
   move r_t3 pc
-  lea r_t3 $\var{ofc}_\var{iter}$
+  lea r_t3 $\var{off}_\var{iter}$
 iter:
@@
 \end{lstlisting}
-Where $\var{ofc}_\var{end}$ and $\var{ofc}_\var{iter}$ are the offsets to the label \texttt{end} % 9
+Where $\var{off}_\var{end}$ and $\var{off}_\var{iter}$ are the offsets to the label \texttt{end} % 9
 and \texttt{iter}, % 2

### 2019-02-12 toplas/paper.tex
[TOPLAS] Update text in discussion and related work section

@@
 
-\emph{Performance and the requirement for stack clearing} The additional
-security measures of the calling convention described in
-\sectionname~\ref{sec:stack-and-return-pointer} impose an overhead compared to a
-calling convention without security guarantees. However, most require only a few
-atomic checks or register clearings on boundary crossings between trusted code
-and adversary, which should produce an acceptable performance overhead. The only
-exception are the requirements for stack clearing that we have in two
-situations: when returning to the adversary and when invoking an adversary
-callback. As we have explained, we need to clear all of the stack that we are
-not using ourselves, not just the part that we have actually used. In other
-words, on every boundary cross between trusted code and adversary code, a
-potentially large region of memory must be cleared.
-
-First, contrary to what we explained before, we actually suspect that this overhead can be avoided in one of the two cases: when returning to the adversary.
+\emph{Performance and the requirement for stack clearing} The additional security measures of the calling convention described in \sectionname~\ref{sec:stack-and-return-pointer} impose an overhead compared to a calling convention without security guarantees.
+However, most require only a few atomic checks or register clearings on boundary crossings between trusted code and adversary, which should produce an acceptable performance overhead.
+The only exception is the stack clearing requirement that we have in two situations: when returning to the adversary and when invoking an adversary callback.
+As we have explained, we need to clear all of the stack that we are not using ourselves not just the part that we have actually used.
+In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be cleared.
+
+First, contrary to what we explained before, we actually suspect that this overhead can be avoided when returning to the adversary.
 In that case, we now think it would suffice to clear only the (much smaller) part of the stack used by the trusted code itself.
@@
 Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenarios where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
-In future work, we are planning to investigate an alternative notion of \emph{linear} capabilities that offer a limited form of revocation without such limitations.
-
-Note, by the way, that CheriBSD's use of local capabilities in CCall does not actually involve a form of revocation.
+In future work, we are planning to investigate an alternative notion of \textit{linear} capabilities that offer a limited form of revocation without such limitations.
+
+CheriBSD's use of local capabilities in CCall does not actually involve a form of revocation.
 CheriBSD's model involves a trusted stack manager that gives every compartment access to its own private stack using a local, write-local capability \citep{Watson2015Cheri}.
@@
 
-\emph{Modularity} It is important that our calling convention is modular, i.e.\
-we do not assume that our code is specially privileged w.r.t.\ the adversary,
-and they can apply the same measures to protect themselves from us as we do to
-protect ourselves from them. More concretely, the requirements we have on
-callbacks and return pointers received from the adversary are also satisfied by
-callbacks and return pointers that we pass to them. For example, our return
-pointers are local capabilities because they must point to memory where we can
-store the old stack pointer, but the adversary's return pointers are also
-allowed to be local. Adversary callbacks are required to be global but the
-callbacks we construct are allocated on the heap and also global.
+\emph{Modularity} It is important that our calling convention is modular, i.e.\ we do not assume that our code is specially privileged w.r.t.\ the adversary, and they can apply the same measures to protect themselves from us as we do to protect ourselves from them.
+More concretely, the requirements we have on callbacks and return pointers received from the adversary are also satisfied by callbacks and return pointers that we pass to them.
+For example, our return pointers are local capabilities because they must point to memory where we can store the old stack pointer, but the adversary's return pointers are also allowed to be local.
+Adversary callbacks are required to be global, but the callbacks we construct are allocated on the heap and also global.
 
@@
 one per thread?
-\emph{Single stack} A single stack is a good choice for the simple capability
-machine presented here, because it works well with higher-order functions. An
-alternative to a single stack would be to have a separate stack per component.
-The trouble with this approach is that, with multiple stacks and local stack
-pointers, it is not clear how components would retrieve their stack pointer upon
-invocation without compromising safety. A safe approach could be to have stack
-pointers stored by a central, trusted stack management component, but it is not
-clear how that could scale to large numbers of separate components. Handling
-large numbers of components is a requirement if we want to use
-capability machines to enforce encapsulation of, for example, every object in an
-object-oriented program or every closure in a functional program.
+\emph{Single stack} A single stack is a good choice for the simple capability machine presented here because it works well with higher-order functions.
+An alternative to a single stack would be to have a separate stack per component.
+The trouble with this approach is that, with multiple stacks and local stack pointers, it is not clear how components would retrieve their stack pointer upon invocation without compromising safety.
+A safe approach could be to have stack pointers stored by a central, trusted stack management component, but it is not clear how that could scale to large numbers of separate components.
+Handling large numbers of components is a requirement if we want to use capability machines to enforce encapsulation of, for example, every object in an object-oriented program or every closure in a functional program.
+
 \\\\
@@
 
-  We do not believe that any of our simplifications are beyond reason.
+  We do not believe that any of our simplifications are beyond reason and expanding the result seems plausible.
   If we added unbounded integers, we would have to change the operational
-  semantics to take over- and under flows into account. This could be achieved
+  semantics to take over and under flows into account. This could be achieved
   by changing the \texttt{plus} and \texttt{minus} instruction to use modulo
@@
   bounds into account. This would require changes to the memory
-  operations (\texttt{store} and \texttt{load}) and the step relation.
+  operations (\texttt{store} and \texttt{load}) and the step relation of the operational semantics.
   The malloc specification would also need to be updated as it would have to
-  fail when it runs out of memory.
+  signal failure when it runs out of memory.
   Finally, expanding the instruction set to a realistic ISA would add a handful
-  of instructions that we would need to reason about.
+  of instructions that we would need to reason about in the proofs.
   All in all, it would add a lot of tedious work to expand this result to a
-  realistic machine. The amount of details in the proofs are already at the
+  realistic machine. The amount of details in the proofs is already at the
   threshold for what should be done with pen and paper, so expanding this work
@@
 programs. A limitation w.r.t. previous work is that the logical relation is
-tailored towards a specific class of problems.
+tailored towards proofs of a specific class of properties.
 
@@
 
-Such properties are hard to prove in our system.
+Such properties are hard to prove in our model.
 For the example, it would be easy to conclude that the returned values are in the value relation (see \figurename~\ref{fig:logrel}).
-This gives us a lot of behavioral information, like conditions under which they are safe to use and invariants that will be preserved when we do, but it does not tell us much about the address of the capability.
-As a very concrete example, capabilities with permission $\noperm$ are always in the value relation irrespective of their address.
+This gives us a lot of behavioral information, like conditions under which the capabilities are safe to use and invariants that will be preserved when we do, but it does not tell us much about the range of authority of the capability.
+As a very concrete example, capabilities with permission $\noperm$ are always in the value relation irrespective of their range of authority.
 Behaviorally, this makes perfect sense; $\noperm$-capabilities cannot be used
@@
 In other situations, however, we may be interested in proving properties like the ones that are often considered in object capability literature: confinement, no authority amplification etc.
-Although such properties are more restrictive and less easy to use for reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation like ours can be adapted to also support them by quantifying the logical relation over a custom interpretation of effectful computations and the type of references.
+Although such properties are more restrictive and tough to use for reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation like ours can be adapted to also support them by quantifying the logical relation over a custom interpretation of effectful computations and the type of references.
 We expect their solution can be readily adapted to our setting modulo some details (like the fact that we do not just have read-write capabilities but also others).
@@
 \noindent\textbf{Logical relation}\\
-\emph{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in
-\figurename~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an
-existing pattern called biorthogonality. Biorthogonality is a pattern for
-defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998}
-in terms of an observation relation of safe configurations (like we do). The
-idea is to define safe evaluation contexts as the set of contexts that produce
-safe observations when plugging safe values and define safe terms as the set of
-terms that can be plugged into safe evaluation contexts to produce safe
-observations. This is an alternative to more direct definitions where safe terms
-are defined as terms that evaluate to safe values. An advantage of
-biorthogonality is that it scales better to languages with control effects like
-call/cc. Our definitions can be seen as a variant of biorthogonality, where we
-take only a single orthogonal closure: we do not define safe evaluation contexts
-but immediately define safe terms as those that produce safe observations when
-plugged with safe values. This is natural because we model arbitrary assembly
-code that does not necessarily respect a particular calling convention: return
-pointers are in principle values like all others and there is no reason to treat
-them specially in the logical relation.
+\emph{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in \figurename~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an existing pattern called biorthogonality.
+Biorthogonality is a pattern for defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998} in terms of an observation relation of safe configurations (like we do).
+The idea is to define safe evaluation contexts as the set of contexts that produce safe observations when plugged with safe values and define safe terms as the set of terms that can be plugged into safe evaluation contexts to produce safe observations.
+This is an alternative to more direct definitions where safe terms are defined as terms that evaluate to safe values.
+An advantage of biorthogonality is that it scales better to languages with control effects like call/cc.
+Our definitions can be seen as a variant of biorthogonality; we take only a single orthogonal closure.
+We do not define safe evaluation contexts but immediately define safe terms as those that produce safe observations when plugged with safe values.
+This is natural because we model arbitrary assembly code that does not necessarily respect a particular calling convention.
+Return pointers are in principle values like all others and there is no reason to treat them specially in the logical relation.
 
@@
 Kripke logical relation for an assembly language (for reasoning about correct
-compilation from ML to assembly), but because they only model non-adversarial
+compilation from ML to assembly). However, because they only model non-adversarial
 code that treats return pointers according to a particular calling convention,
@@
   \emph{Why use a logical relation and not a simpler proof technique?}
-  The concept of a capability exists on different levels of the computer stack, so capabilities have been studied and safety properties proven about them in other contexts than assembly languages.
+  The concept of a capability exists on different levels of abstraction on computers, so capabilities have been studied and safety properties proven about them in other contexts than assembly languages.
   Traditionally, logical relations have not been used for safety property proofs, so why do we need one here?
@@
   The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject a can have a write capability for memory B.}
-  However, their policy are, in a sense, binary: either subject a can not write to memory B, or it can write anything.
-  Our approach is much more fine-grained, allowing us to define, for example, a policy that subject a can write to memory B, but only if the value being written is an even number.
+  However, their policy are, in a sense, binary: subject a can either not write or write anything to memory B.
+  Our approach is much more fine-grained and allows us to define, for example, a policy that subject a can write to memory B, but only if the value is an even number.
   Because our invariants are indexed by worlds themselves, we can even define a policy that capabilities written to memory B must respect certain other invariants themselves.
   It is exactly the expressive invariants that enable us to prove the awkward example.
-  We place a protocol on the memory where variable \texttt{x} is stored that says that it sometimes can be either $0$ or $1$, and at other times it has to be $1$.
+  We place a protocol on the memory where variable \texttt{x} is stored that says that it sometimes can be either $0$ or $1$ and at other times it has to be $1$.
 
@@
 \label{sec:related-work}
-
-Finally, we summarize how our work relates to previous work. We do not
-repeat the work we discussed in \sectionname~\ref{sec:discussion}.
-
-Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to \citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview
-of previous work. The capability machine formalized in
-\sectionname~\ref{sec:capab-mach-with} is a simple but representative
-model modeled mainly after the
-M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers
-resemble the M-Machine's) and
-CHERI~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the
-memory and local capabilities resemble CHERI's). The latter is a
-recent and relatively mature capability machine, which combines
-capabilities with a virtual memory approach, in the interest of
-backwards compatibility and gradual adoption. As discussed, our local
-capabilities can cross module boundaries, 
-contrary to what is enforced by CHERI's default CCall implementation.
-
-Plenty of other papers\lau{If there are plenty, then I guess we should cite more
-  than one?} enforce well-bracketed control flow at a low level, but most are
-restricted to preventing particular types of attacks and enforce only partial
-correctness of control flow. This includes particularly the line of work on
-\emph{control-flow integrity}~\citep{abadi_control-flow_2005}. Those use a quite
-different attacker model than us: they assume an attacker that is not able to
-execute code, but can overwrite arbitrary data at any time during execution (to
-model buffer overflows). By checking the address of every indirect jump and
-using memory access control to prevent overwriting code, this work enforces what
-they call control-flow integrity, formalized as the property that every jump
-will follow a legal path in the control-flow graph. As discussed in
-\sectionname~\ref{sec:discussion}, such a property ignores temporal properties and
-seems hard to use for reasoning.
+In this section, we summarize how our work relates to previous work.
+We do not repeat the work we discussed in \sectionname~\ref{sec:discussion}.
+
+Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to \citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview of previous work.
+The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is a simple but representative model modeled mainly after the M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers resemble the M-Machine's) and CHERI~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the memory and local capabilities resemble CHERI's).
+The latter is a recent and relatively mature capability machine. CHERI combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
+As discussed, our local capabilities can cross module boundaries contrary to what is enforced by CHERI's default CCall implementation.
+
+Plenty of other papers\lau{If there are plenty, then I guess we should cite more than one?}
+enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
+This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
+This line of work uses a quite different attacker model than us.
+They assume an attacker that is not able to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
+By checking the address of every indirect jump and using memory access control to prevent overwriting code, this work enforces what they call control-flow integrity formalized as the property that every jump will follow a legal path in the control-flow graph.
+As discussed in \sectionname~\ref{sec:discussion}, such a property ignores temporal properties and seems hard to use for reasoning.
 
@@
 primitive (a capability machine with local capabilities rather than a machine
-with a primitive notion of compartments) and we do not use a trusted stack
-manager, but a decentralized calling convention based on local
+with a primitive notion of compartments). Further, we do not use a trusted stack
+manager but a decentralized calling convention based on local
 capabilities. Also, both prove a secure compilation result from a high-level
-language, which clearly implies a general form of control-flow correctness,
+language which clearly implies a general form of control-flow correctness
 while we define a logical relation that can be used to reason about
@@
 their notion of effect parametricity. Our logical relation uses several ideas
-from previous work, like Kripke worlds with regions containing state transition
+from previous work like Kripke worlds with regions containing state transition
 systems \citep{Ahmed:popl09}, public/private future worlds \citep{Dreyer:jfp12}
@@
 Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
-separation logic and is formalized in Coq, building on the Iris Proof Mode for
+separation logic, and is formalized in Coq building on the Iris Proof Mode for
 Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
@@
 \begin{toplas}
-In the future, we would also like to investigate a new program logic for
+In the future, we would like to develop a new program logic for
 reasoning about capability safety for our capability machine model. In fact,
@@
 framework that can be instantiated with different programming languages. OCPL
-was able to leverage existing Iris specifications for a high-level language;
+was able to leverage existing Iris specifications for a high-level programming language;
 for our capability machine model, however, it would be necessary to devise new
@@
 If we want to scale this approach even further, we would like to reason at a higher level of abstraction namely at the level of a high-level language.
-That is, we would rather construct a program logic for a high-level language, so we can reason about the programs the language we actually write them in and at the same time get the guarantees for the programs we are going to execute, i.e.\ the compiled programs.
+That is, we would rather construct a program logic for a high-level programming language, so we can reason about the programs in the language we actually write them in and at the same time get guarantees about the compiled program.
 To achieve this, we would have to construct a secure compilation \citep{patrignani_secure_compilation_2019} that preserves the security abstraction of the high-level language.
-The experiences made about logical relations for capability machines here has been used by \citet{skorstengaard_stktokens_2019} to prove make advances in the direction of building a secure compilation targeting capability machines.
-They present a calling convention based on linear capabilities rather than local capabilities, and they provide a fully abstract compilation that can be used as the last phase in a secure compilation.
+The experiences made about logical relations for capability machines here has been used by \citet{skorstengaard_stktokens_2019} to make advances in the direction of building a secure compilation targeting capability machines.
+They present a calling convention based on linear capabilities rather than local capabilities, and they provide a fully-abstract compilation that can be used as the last phase in a secure compilation.
 The full-abstraction proof uses a logical relation.

### 2019-02-13 toplas/paper.tex
[TOPLAS] Consolidated POPL19 reference into one paragraph

@@
 \noindent\textbf{Calling convention}\\
-\emph{Formulating control flow correctness} While we claim that our calling
-convention enforces control-flow correctness, we do not prove a general theorem
-that shows this because it is not clear what such a theorem should look like.
+\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
 Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
   Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
-In fact, our examples show that our logical relation implies a stronger form of
-control-flow correctness than such formulations, although this is not made very
-explicit. As future work, we consider looking at a more explicit and useful way
-to formalize control-flow correctness. The idea would be to define a variant
-of our capability machine with call and return instructions and
-well-bracketed control flow built-in to the operational semantics, and then
-prove that compiling such code to our machine using our calling convention is
-fully abstract~\citep{abadi_protection_1998}.
+In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
+In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.
+The idea of their definition is to define a variant of a simple capability machine with a built-in stack as well as call and return instructions.
+The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
+They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to the one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
+Their full abstraction proof uses a logical relation, but it differs in a number of ways.
+First, their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
+This means that it could not be used to prove program correctness results similar to ours.
+Their calling convention is based on something called linear capabilities that offer a different kind of limited revocation.
 \itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
@@
 Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenarios where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
-In future work, we are planning to investigate an alternative notion of \textit{linear} capabilities that offer a limited form of revocation without such limitations.
 
@@
 Handling large numbers of components is a requirement if we want to use capability machines to enforce encapsulation of, for example, every object in an object-oriented program or every closure in a functional program.
-
 \\\\
@@
 To achieve this, we would have to construct a secure compilation \citep{patrignani_secure_compilation_2019} that preserves the security abstraction of the high-level language.
-The experiences made about logical relations for capability machines here has been used by \citet{skorstengaard_stktokens_2019} to make advances in the direction of building a secure compilation targeting capability machines.
-They present a calling convention based on linear capabilities rather than local capabilities, and they provide a fully-abstract compilation that can be used as the last phase in a secure compilation.
-The full-abstraction proof uses a logical relation.
 \end{toplas}

### 2019-02-14 toplas/paper.tex
[TOPLAS] Rewrite StkTokens reference slightly

@@
 The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
-They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to the one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
-Their full abstraction proof uses a logical relation, but it differs in a number of ways.
-First, their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
+They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
+Their full abstraction proof uses a logical relation, but it differs in a number of ways from the one we present.
+Their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
 This means that it could not be used to prove program correctness results similar to ours.
-Their calling convention is based on something called linear capabilities that offer a different kind of limited revocation.
+Their calling convention is based on something called linear capabilities rather than local capabilities.
+Linear capabilities offer a different kind of limited revocation.
 \itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}

### 2019-02-25 toplas/paper.tex
[TOPLAS] Update syntax figure to be less dense
[TOPLAS] Rephrase local capability explanation

@@
 \begin{abstract}
-  Capability machines provide security guarantees at machine level which makes
-  them an interesting target for secure compilation schemes that provably
-  enforce properties such as control-flow correctness and encapsulation of local
-  state. We provide a formalization of a representative capability machine with
-  local capabilities and study a novel calling convention.  We provide a logical
-  relation that semantically captures the guarantees provided by the hardware (a
-  form of capability safety) and use it to prove control-flow correctness and
-  encapsulation of local state.  The logical relation is not specific to our
-  calling convention and can be used to reason about arbitrary programs.
+  Capability machines provide security guarantees at machine level which makes them an interesting target for secure compilation schemes that provably enforce properties such as control-flow correctness and encapsulation of local state.
+  We provide a formalization of a representative capability machine with local capabilities and study a novel calling convention.
+  We provide a logical relation that semantically captures the guarantees provided by the hardware (a form of capability safety) and use it to prove control-flow correctness and encapsulation of local state.
+  The logical relation is not specific to our calling convention and can be used to reason about arbitrary programs.
 \end{abstract}
@@
 \[\arraycolsep=1.4pt
-  \begin{array}{rlrcl l rlrcl}
-    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
-    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
-    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
-    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
-    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
-    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
-     && \Caps  &\isdef & \multicolumn{6}{l}{\left\{
+  \begin{array}{rlrcl}
+    \addr   &\in& \Addrs &\isdef& \nats \\
+    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid\exec\mid \entry\mid \rwx\mid \rwlx  \\
+    \gl&\in&\Globals & ::= &\glob \mid \local \\
+    c &\in& \Caps  &\isdef & \left\{
                          ((\perm,\gl),\start,\addrend,\addr)\mid
                          \start,\addr\in\Addrs,\addrend \in
-                         \Addrs\cup \{\infty\}\right\}}
+                         \Addrs\cup \{\infty\}\right\}\\
+    w &\in&\Words &\isdef& \ints + \Caps \\
+    r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
+    \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
+    m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
+    \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
+    \ms  &\in & \MemSegments &\isdef& \Addrs \parfun \Words \\
+    &&\Confs &\isdef& \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps
   \end{array}
@@
 
-Finally, the $\store{}{}$ instruction updates the memory to the requested value
-if the capability has write authority for the requested location. However, the
+Finally, the $\store{}{}$ instruction updates the memory to the argument value
+if the capability has write authority for the specified location. However, the
 instruction is complicated by the presence of \emph{local capabilities} modeled
-after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
-capabilities are special in that they can only be kept in registers, i.e.\ they
+after the ones in the CHERI processor~\citep{Watson2015Cheri}. At a high-level, local
+capabilities are special in that they can only be kept in registers, i.e.\, they
 cannot be stored to memory. This means that local capabilities can be
-\emph{temporarily} given to an adversary, for the duration of an invocation: if
+\emph{temporarily} given to an adversary, for the duration of an invocation. If
 we make sure to clear the capability from the register file after control is
-passed back to us, they will not have been able to store the capability.
+passed back to us, the adversary is unable to store the capability.
 However, there is one exception to the rule above: local capabilities can be
 stored to memory for which we have a capability with write-local authority
-(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
+(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack
 where register contents can be stored, including local capabilities. As long as
@@
 This line of work uses a quite different attacker model than us.
-They assume an attacker that is not able to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
+They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
 By checking the address of every indirect jump and using memory access control to prevent overwriting code, this work enforces what they call control-flow integrity formalized as the property that every jump will follow a legal path in the control-flow graph.

### 2019-02-25 presentations/aarhus-25-02-2019/presentation.tex
Add Aarhus February presentation

--- 2019-01-24 presentations/popl19-01-2019/presentation.tex
@@
 \begin{document}
+\beamertemplatenavigationsymbolsempty
 \maketitle
@@
     capability aware.}
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Well-bracketed control flow and local state encapsulation}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+\begin{lstlisting}[language=C, escapechar=!]
+!\tikz[remember picture] \node[] (a) {};!void a()       !\tikz[remember picture] \node[] (bt) {};! 
+ {
+   !\tikz[remember picture] \node[] (pc6) {};!...
+   !\tikz[remember picture] \node[] (pc7) {};!return;!\tikz[remember picture] \node[] (ret) {};!      
+ }              !\tikz[remember picture] \node[] (bb) {};!
+
+ void b()
+ {
+   !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
+   !\tikz[remember picture] \node[] (pc2) {};!a();
+   !\tikz[remember picture] \node[] (pc3) {};!...
+   !\tikz[remember picture] \node[] (pc4) {};!a();
+   !\tikz[remember picture] \node[] (pc5) {};!return;   
+ }
+\end{lstlisting}
+% Mention that our work support high-order
+% No good definition. We appeal to you intuitive definition
+      \begin{tikzpicture}[remember picture, overlay]
+%        \draw<1> node { };
+        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=0.2cm] (brace) [midway] {};
+        \draw<2> node[right = -0cm of brace, text width = 3cm] {Function \texttt{a} cannot access variable \texttt{x}};
+
+        \draw<3> ($(pc4) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc4);
+        \draw<4> ($(a) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (a);
+        \draw<5> ($(pc6) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc6);
+        \draw<6> ($(pc7) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc7);
+        \draw<7-> node[below right = of ret] (w) {\huge ?};
+        \draw<7-> (ret) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202},bend left] (w);
+        \draw<8> ($(pc5) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc5);
+        \draw<9> ($(pc3) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc3);
+      \end{tikzpicture}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \only<2>{Local-state encapsulation (LSE)}
+      \only<3->{Well-bracketed control flow (WBCF)}
+    \end{column}
+    % What is LSE and WBCF
+    % Quick example?
+  \end{columns}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Attack on naive stack and return capabilities }
+  \frametitle{Attack 1 on naive stack and return capabilities }
   \centering
@@
 \begin{frame}
-  \frametitle{\stktokens{} prevent the attack}
+  \frametitle{\stktokens{}}
+  \begin{itemize}
+  \item Check the base address of the stack capability before and after calls.
+  \item Make sure that local stack frames are non-empty.
+    % Not illustrated. Trusted code must claim spot on the call stack to make
+    % sure that they are returned to.
+  \item Create token and data return capability on call: split the stack
+    capability in two to get a stack capability for your local stack frame and a
+    stack capability for the unused part of the stack. The former is sealed and
+    used for the data part of the return pair.
+  \item Create code return capability on call: Seal the old program pointer.
+  \item Reasonable use of seals: Return seals are only used to seal old program
+    pointers, every return seal is only used for one call site, and they are not
+    leaked. % If return seals are used for other things, like sealing arbitrary
+            % code pointers, then they can be used as an unsealing mechanism.
+            % If the same seal is used for multiple call sites, then we lose the
+            % guarantee that we return to the correct place. One could use the
+            % wrong return capability to return "out of order"
+            % If return seals are leaked, then we have no guarantee about what
+            % they are used for.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} prevent attack 1}
   % Emphazise that it is dynamically enforced.
@@
 
+\begin{frame}
+  \frametitle{Attack 2 on naive stack and return capabilities }
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    % first step: capabilities
+    \begin{onlyenv}<+>
+      \draw[fill=white] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize current stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
+      % \draw node[fill=gray!50,right=0cm of oldframe1] (rp1l) {return pointer data};
+    \end{onlyenv}
+
+    % attacker stores copy of stack pointer high in the stack
+    \draw<+->[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
+    \begin{onlyenv}<.-.(2)>
+      \actadv{(0,4)}{(6,7)}{callee stack frame}
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,4) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+    \draw<+->[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=myred] {\footnotesize copy of sp};
+    \draw<.> (6,13.25) edge[myred,thick,->,in=45] (sp1);
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
+      \draw (6,13.25) edge[myred,thick,->,in=45] (spold);
+    \end{onlyenv}
+
+    % attacker calls trusted code again
+%    \draw<+->[fill=gray!50] (0,4) rectangle (6,7) node[pos=.5] {\footnotesize callee stack frame};
+    \begin{onlyenv}<+->
+      \inactadv{(0,4)}{(6,7)}{callee stack frame}
+    \end{onlyenv}
+    \begin{onlyenv}<.>
+      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+
+    % attacker uses old stack pointer
+    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
+    \begin{onlyenv}<.-.(1)>
+      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
+      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw node[right=0cm of sp1] { stack pointer };
+    \end{onlyenv}
+    \draw<+> (spold) edge[->,color=myred,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} prevent attack 2}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+
+    \draw (16.5,5.5) node {register file};
+    \begin{scope}
+      \clip (15.9,1.1) rectangle (17.1,5.1);
+      \foreach \x in {1,2,...,4}
+      {
+        \draw[fill=white] (16,\x) rectangle (17,\x+1) node[pos=.5,color=mygreen] {};
+      };
+    \end{scope}
+    \node (r1) at (16.75,4.5) {};
+    \node (r2) at (16.75,3.6) {};
+
+    % Caller 
+    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+    \begin{onlyenv}<+>
+      \capbrace{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw (r1) edge[->,myred,thick,bend left,dashed] (sp1);
+    \end{onlyenv}
+
+    % Split the stack pointer
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
+      \draw (r2) edge[->,myred,thick,bend left,dashed] (sp2);
+      \draw (r1) edge[->,myred,thick,bend left, dashed] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    % Callee called
+    \begin{onlyenv}<+>
+      \capbrace[usp]{(6.5,5)}{(6.5,16)}
+      \draw (r1) edge[->,myred,bend left,dashed] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+
+    \begin{onlyenv}<.->
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[fill=gray!50,right=0cm of sp2] { \phantom{return data pointer} };
+      \draw (r2) edge[->,myred,bend left,dashed] (sp2);
+      \draw node[right=0cm of sp2] { return data pointer };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+>
+
+      \capbrace[usp]{(13.5,5)}{(13.5,16)}
+      \draw (6,13.25) edge[myred,thick,->,in=45,dashed] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+
+      \draw[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=myred] {\footnotesize sp};
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actadv{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+
+    % At this point we have a linear point stored in a part of memory that it governs.
+In other words, no other capabilities can grant access to this part of memory, to the capabilitiy is essentially lost.
+%    \draw<+> node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (4,10) { Stack pointer\\essentially lost.};
+
+    \begin{onlyenv}<+->
+
+      \capbrace[usp]{(13.5,5)}{(13.5,16)}
+      \draw (6,13.25) edge[myred,thick,->,in=45,dashed] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+
+      \begin{scope}
+        \fill[fill=gray!50] (0,2.5) rectangle (6,15) node[pos=.5,color=black] {};
+        \draw (0,0) -- (0,15);
+        \draw (6,0) -- (6,15);
+      \end{scope}
+      \draw[fill=gray!50] (0,13) rectangle (6,13.5) node[pos=.5,color=myred,dashed] {\footnotesize sp};
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \inactadv{(0,5)}{(6,8)}{callee stack frame}
+      \actsf{(0,8)}{(6,10.5)}{callee (2) stack frame} 
+  \end{onlyenv}
+
+  \draw<+>[mygreen,very thick] (-0.5,7.5) -- (6.5,11);
+  \draw<.>[mygreen,very thick] (-0.5,11) -- (6.5,7.5);
+  \draw<.> node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (14,7) { No stack pointer\\for callee (2)!};
+
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \frametitle{Attack 3 on naive stack and return capabilities}
+    \begin{scope}
+    \clip (-.1,-.1) rectangle (5.1,10.1);
+    \fill[fill=gray!50] (0,0) rectangle (5,10);
+    \draw (0,0) -- (0,10);
+    \draw (5,0) -- (5,10);
+    \end{scope}
+    \draw (2.5,11) node {code memory};
+
+    \begin{scope}
+    \clip (12.9,3.9) rectangle (18.1,11.1);
+    \fill[fill=gray!50] (13,4) rectangle (18,11);
+    \draw (13,4) -- (13,11);
+    \draw (18,4) -- (18,11);
+    \end{scope}
+    \draw (15.5,12) node {more code memory};
+
+    \begin{scope}
+    \clip (12.9,-.1) rectangle (18.1,2.1);
+    \fill[fill=gray!50] (13,0) rectangle (18,2);
+    \draw (13,0) -- (13,2);
+    \draw (18,0) -- (18,2);
+    \end{scope}
+    \draw (14.5,3) node {heap memory};
+
+\note{Caller calls callee}
+    \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+
+    \begin{onlyenv}<+->
+      \capbrace[ret1]{(5.1,1)}{(5.1,4)}
+      \draw node[fill=gray!50, right=0cm of ret1] { \phantom{ret-ptr} };
+      \draw node[right=0cm of ret1] { ret-ptr };
+    \end{onlyenv}
+
+\note{Callee has access to some place to store things.}
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=white] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
+    \end{onlyenv}
+
+    \begin{onlyenv}<.-.(2)>
+      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
+      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
+      \draw (18,5.5)  edge[->,bend left] (copy);
+    \end{onlyenv}
+
+\note{Callee stores the return pointer they got from caller.}
+    \begin{onlyenv}<+->
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{myred} copy of ret-ptr}};
+      \draw (13,1.25) edge[->,myred,thick,bend left] (ret1);
+    \end{onlyenv}
+
+\note{Callee calls callee (1)}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=3pt,raise=4pt},yshift=0pt]
+      (12.9,5) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
+      \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
+      \draw node[left=0cm of ret2] (retptr1) { ret-ptr (1) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=gray!30] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{myred} copy of ret-ptr}};
+    \end{onlyenv}
+
+\note{Callee (1) calls callee (2)}
+    \begin{onlyenv}<+->
+      \capbrace[ret3]{(5.1,5)}{(5.1,8)}
+      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
+      \draw node[right=0cm of ret3] (retptr2) { ret-ptr (2) };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
+      \draw[fill=white] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
+    \end{onlyenv}
+\note{Callee (2) colludes with callee}
+    \begin{onlyenv}<+->
+      \draw (21,7.5) node[devil,opaque=0.5,mirrored,minimum size=1cm] {};
+    \end{onlyenv}
+\note{Callee (2) has access to ret-ptr (the return pointer caller passed to callee)}
+    \begin{onlyenv}<+->
+      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
+      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
+      \draw (18,8.25)  edge[->,myred,thick,bend left] (copy);
+      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{myred} copy of ret-ptr}};
+    \end{onlyenv}
+
+\note{Callee (2) uses ret-ptr to return}
+    \begin{onlyenv}<+->
+      \draw[fill=gray!30] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
+      \draw[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
+      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{myred} copy of ret-ptr}};
+    \end{onlyenv}
+
+    \note{This breaks well-bracketedness as ret-ptr (2) and ret-ptr (1) was not returned to}
+    \begin{onlyenv}<+->
+      \draw node [ellipse, fit=(retptr1) (retptr2), draw, myred] { };
+      \draw node[myred,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  
+    \end{onlyenv}
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} prevent attack 3}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+  \scope
+    \clip (-.1,-.1) rectangle (6.1,15.1);
+    \fill[fill=white] (0,0) rectangle (6,15);
+    \draw (0,0) -- (0,15);
+    \draw (6,0) -- (6,15);
+    \draw[fill=gray!50] (0,-.5) rectangle (6,1.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+  \endscope
+  % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+  \begin{scope}
+    \fill[gray!50] (13,0) rectangle (18,13);
+    \draw (13,0) -- (13,13);
+    \draw (18,0) -- (18,13);
+    \draw (15.5,14) node {code memory};
+  \end{scope}
+
+\note{Caller calls callee (split stack pointer to make ret-ptr-d)}
+  \begin{onlyenv}<+->
+    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \draw[fill=white] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \capbrace[sp]{(6.1,1.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \capbrace[ret1]{(6.1,1.5)}{(6.1,4.5)}
+    \draw node[fill=gray!50,right=0cm of ret1] { \phantom{ret-ptr-d} };
+    \draw node[right=0cm of ret1] { ret-ptr-d };
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \draw [decorate,decoration={brace,aspect=0.3,amplitude=4pt,raise=4pt},yshift=0pt]
+    (12.9,1) -- (12.9,3) node[draw=black] (ret1c) [black,pos=0.3,xshift=-0.6cm] {};
+    \draw node[fill=gray!50,left=0cm of ret1c] { \phantom{ret-ptr-c} };
+    \draw node[left=0cm of ret1c] { ret-ptr-c };
+  \end{onlyenv}
+
+  \begin{onlyenv}<.-.(3)>
+    \capbrace[sp]{(6.1,4.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+-+(3)>
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=white] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+  \end{onlyenv}
+
+\note{Callee saves ret-ptr-d on the heap.}
+  \begin{onlyenv}<+->
+    \draw (13,4.75) edge[myred,bend right,thick,->,dashed] (ret1);
+  \end{onlyenv}
+\note{Callee saves ret-ptr-c on the heap.}
+  \begin{onlyenv}<+->
+    \draw (13,4.75) edge[myred,bend right,thick,->] (ret1c);
+  \end{onlyenv}
+
+\note{Callee calls callee (1)}
+\note{Before the callee jumps, it is assumed that it has checked the global "max height" of its stack pointer!}
+  \begin{onlyenv}<+-+(4)>
+    \capbrace[ret2]{(6.1,4.5)}{(6.1,7.5)}
+    \draw node[fill=gray!50,right=0cm of ret2] { \phantom{ret-ptr-d (1)} };
+    \draw node[right=0cm of ret2] { ret-ptr-d (1) };
+  \end{onlyenv}
+  \begin{onlyenv}<.-.(1)>
+    \capbrace[sp]{(6.1,7.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+-+(1)>
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \actsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \draw[fill=white] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+  \end{onlyenv}
+
+\note{Callee (1) calls callee (2)}
+  \begin{onlyenv}<+-+(2)>
+    \capbrace[ret3]{(6.1,7.5)}{(6.1,10.5)}
+    \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr-d (2)} };
+    \draw node[right=0cm of ret3] { ret-ptr-d (2) };
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \capbrace[sp]{(6.1,10.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \inactsf{(0,7.5)}{(6,10.5)}{callee stack frame (1)}
+  \end{onlyenv}
+
+  \begin{onlyenv}<.->
+    \draw[fill=gray!30] (13,1) rectangle (18,3) node[pos=.5] {\footnotesize caller code};
+    \draw[fill=gray!30] (13,4) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
+    \draw[fill=gray!30] (13,7) rectangle (18,9) node[pos=.5] {\footnotesize callee code (1)};
+    \draw[fill=white] (13,10) rectangle (18,12) node[pos=.5] {\footnotesize callee code (2)};
+  \end{onlyenv}
+  \begin{onlyenv}<.-.(1)>
+    \actadv{(0,10.5)}{(6,13.5)}{callee stack frame (2)}
+  \end{onlyenv}
+
+\note{Callee (2) has a capability for Callee and can thus access the stored ret-ptr-d}
+  \begin{onlyenv}<+->
+    \draw [decorate,decoration={brace,amplitude=4pt,mirror,raise=4pt},yshift=0pt]
+    (18.1,4) -- (18.1,6) node[draw=black] (calleecap) [black,midway,xshift=0.6cm] {};
+    \draw (18,11) edge[myred,bend left,thick,->] (calleecap);
+  \end{onlyenv}
+
+\note{Callee (2) uses ret-ptr-d to return}
+  \begin{onlyenv}<+->
+    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \draw node[fill=white,right=0cm of ret1] { ret-ptr-d };
+    \draw node[fill=white,left=0cm of ret1c] { ret-ptr-c };
+    \draw (13,4.75) edge[myred,bend right,thick,->,dashed] (ret1);
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \draw[mygreen,very thick] (6.4,4.4) -- (7.4,5.4);
+    \draw[mygreen,very thick] (7.4,4.4) -- (6.4,5.4);
+    \draw node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (10,6) { Splice fails!};  
+  \end{onlyenv}
+
+\note{Splicing: fails because ends of capabilities don't meet. callee stack frame (1) cannot be empty (it will at least contain 42).}
+  % \begin{onlyenv}<+->
+  %   \draw node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (8,7.5) {Caller tries to splice sp \\and ret-ptr-d, but fails!};  
+  % \end{onlyenv}
+\note{Before the caller makes another call/return, it will check that the stack height matches the global max stack height.}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} enforces single stack}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    % recurrent parts
+  \scope
+    \clip (-.1,-.1) rectangle (6.1,15.1);
+    \fill[fill=white] (0,0) rectangle (6,15);
+    \draw (0,0) -- (0,15);
+    \draw (6,0) -- (6,15);
+    \draw[fill=gray!50] (0,-.5) rectangle (6,1.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+  \endscope
+  % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+  \begin{scope}
+    \fill[gray!50] (13,0) rectangle (18,13);
+    \draw (13,0) -- (13,13);
+    \draw (18,0) -- (18,13);
+    \draw (15.5,14) node {heap};
+  \end{scope}
+
+\note{Caller calls callee (split stack pointer to make ret-ptr-d)}
+  \begin{onlyenv}<+>
+    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \capbrace[sp]{(6.1,1.5)}{(6.1,15)}
+    \draw node[right=0cm of sp] { sp };
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \draw[fill=white] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \capbrace[sp2]{(6.1,4.5)}{(6.1,15)}
+    \capbrace[sp]{(6.1,1.5)}{(6.1,4.5)}
+    \draw node[right=0cm of sp2] { sp2 };
+    \draw node[right=0cm of sp, fill=gray!50] { ret-ptr-d };
+    \draw (13,6.5) edge[myred,bend right,thick,->,dashed] (sp);
+  \end{onlyenv}
+
+  \begin{onlyenv}<+>
+    \capbrace[sp3]{(6.1,9.5)}{(6.1,15)}
+    \capbrace[sp2]{(6.1,4.5)}{(6.1,9.5)}
+    \capbrace[sp]{(6.1,1.5)}{(6.1,4.5)}
+    \draw node[right=0cm of sp3] { sp3 };
+    \draw node[right=0cm of sp2] { sp2 };
+    \draw node[right=0cm of sp, fill=gray!50] { ret-ptr-d };
+    \draw (13,7.5) edge[myred,bend right,thick,->,dashed] (sp2);
+    \draw (13,6.5) edge[myred,bend right,thick,->,dashed] (sp);
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \actsf{(0,9.5)}{(6,11.5)}{callee stack frame (1)}
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \draw[fill=gray!30] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+  \end{onlyenv}
+
+  \begin{onlyenv}<+>
+    \capbrace[sp4]{(6.1,11.5)}{(6.1,15)}
+    \capbrace[sp3]{(6.1,9.5)}{(6.1,11.5)}
+    \capbrace[sp2]{(6.1,4.5)}{(6.1,9.5)}
+    \capbrace[sp]{(6.1,1.5)}{(6.1,4.5)}
+    \draw node[right=0cm of sp4] { sp4 };
+    \draw node[right=0cm of sp3, fill=gray!50] { sp3 };
+    \draw node[right=0cm of sp2] { sp2 };
+    \draw node[right=0cm of sp, fill=gray!50] { ret-ptr-d };
+    \draw (13,8.5) edge[myred,bend right,thick,->,dashed] (sp3);
+    \draw (13,7.5) edge[myred,bend right,thick,->,dashed] (sp2);
+    \draw (13,6.5) edge[myred,bend right,thick,->,dashed] (sp);
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \inactsf{(0,9.5)}{(6,11.5)}{callee stack frame (1)}
+    \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+  \end{onlyenv}
+  \begin{onlyenv}<.>
+    \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+    \draw[fill=white] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+  \end{onlyenv}
+
+  \begin{onlyenv}<+->
+    \capbrace[sp4]{(6.1,11.5)}{(6.1,15)}
+    \capbrace[sp3]{(6.1,9.5)}{(6.1,11.5)}
+    \capbrace[sp]{(6.1,1.5)}{(6.1,9.5)}
+    \draw node[right=0cm of sp4, fill=gray!50] { sp4 };
+    \draw node[right=0cm of sp3, fill=gray!50] { sp3 };
+    \draw node[right=0cm of sp] { sp };
+    \draw (13,8.5) edge[myred,bend right,thick,->,dashed] (sp3);
+  \end{onlyenv}
+  \begin{onlyenv}<.->
+    \inactsf{(0,9.5)}{(6,11.5)}{callee stack frame (1)}
+    \inactadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+    \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+    \draw[fill=white] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+  \end{onlyenv}
+  \begin{onlyenv}<+->
+    \draw[mygreen,very thick] (2.5,10) -- (3.5,9);
+    \draw[mygreen,very thick] (3.5,10) -- (2.5,9);
+    \draw node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center, anchor=west] () at (4,9.5) { Base check fails!};  
+  \end{onlyenv}
+\end{tikzpicture}
+\end{frame}
+
 % \begin{frame}
@@
 
+% \begin{frame}
+%   \frametitle{\stktokens{}: ``Base address'' check}
+%   \includegraphics[width=\textwidth]{stk-fig1}
+% \end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} summary}
+  \setbeamercovered{transparent}
+  \begin{itemize}
+  \item<0> Check the base address of the stack capability before and after calls.
+  \item<1> Make sure that local stack frames are non-empty.
+    % Not illustrated. Trusted code must claim spot on the call stack to make
+    % sure that they are returned to.
+  \item<0> Create token and data return capability on call: split the stack
+    capability in two to get a stack capability for your local stack frame and a
+    stack capability for the unused part of the stack. The former is sealed and
+    used for the data part of the return pair.
+  \item<0> Create code return capability on call: Seal the old program pointer.
+  \item<2> Reasonable use of seals: Return seals are only used to seal old program
+    pointers, every return seal is only used for one call site, and they are not
+    leaked. % If return seals are used for other things, like sealing arbitrary
+            % code pointers, then they can be used as an unsealing mechanism.
+            % If the same seal is used for multiple call sites, then we lose the
+            % guarantee that we return to the correct place. One could use the
+            % wrong return capability to return "out of order"
+            % If return seals are leaked, then we have no guarantee about what
+            % they are used for.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} FAQ}
+  \begin{itemize}
+  \item \textit{Do you support tail calls?}
+    \begin{itemize}
+    \item Yes.
+    \end{itemize}
+  \item \textit{Do you support higher-order functions?}
+    \begin{itemize}
+    \item Yes.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \section{Defining well-bracketed control flow and local state encapsulation by fully-abstract overlay semantics}
-\begin{frame}[fragile]
-  \frametitle{Well-bracketed control flow and local state encapsulation}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-\begin{lstlisting}[language=C, escapechar=!]
-!\tikz[remember picture] \node[] (a) {};!void a()       !\tikz[remember picture] \node[] (bt) {};! 
- {
-   !\tikz[remember picture] \node[] (pc6) {};!...
-   !\tikz[remember picture] \node[] (pc7) {};!return;!\tikz[remember picture] \node[] (ret) {};!      
- }              !\tikz[remember picture] \node[] (bb) {};!
-
- void b()
- {
-   !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
-   !\tikz[remember picture] \node[] (pc2) {};!a();
-   !\tikz[remember picture] \node[] (pc3) {};!...
-   !\tikz[remember picture] \node[] (pc4) {};!a();
-   !\tikz[remember picture] \node[] (pc5) {};!return;   
- }
-\end{lstlisting}
-% Mention that our work support high-order
-% No good definition. We appeal to you intuitive definition
-      \begin{tikzpicture}[remember picture, overlay]
-%        \draw<1> node { };
-        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=0.2cm] (brace) [midway] {};
-        \draw<2> node[right = -0cm of brace, text width = 3cm] {Function \texttt{a} cannot access variable \texttt{x}};
-
-        \draw<3> ($(pc4) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc4);
-        \draw<4> ($(a) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (a);
-        \draw<5> ($(pc6) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc6);
-        \draw<6> ($(pc7) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc7);
-        \draw<7-> node[below right = of ret] (w) {\huge ?};
-        \draw<7-> (ret) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202},bend left] (w);
-        \draw<8> ($(pc5) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc5);
-        \draw<9> ($(pc3) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc3);
-      \end{tikzpicture}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \only<2>{Local-state encapsulation (LSE)}
-      \only<3->{Well-bracketed control flow (WBCF)}
-    \end{column}
-    % What is LSE and WBCF
-    % Quick example?
-  \end{columns}
-\end{frame}
 
@@
 
-% \begin{frame}
-%   \frametitle{Fully-abstract overlay semantics}
-%   \begin{columns}
-% \begin{column}{0.5\textwidth}
-%   \begin{itemize}[<+->]
-%   \item Compile from oLCM to LCM
-%   \item Compilation is the identity
-%   \item Builtin call stack abstraction in oLCM preserved after compilation
-%   \end{itemize}
-% \end{column}
-% \begin{column}{0.5\textwidth}
-%   \begin{tikzpicture}
-%     \draw (0,0) node[draw=black, rounded corners] (olcm) {oLCM};
-%     \draw node[draw=black, rounded corners,below=2.5cm of olcm] (lcm) {LCM};
-%     \draw (olcm) edge[bend left, ->,thick] (lcm);
-%     \draw node[left = 3cm of olcm] (ll1) {};
-%     \draw node[right = 3cm of olcm] (rl1) {};
-%     \draw node[below = 1.25cm of ll1] (ll) {};
-%     \draw node[below = 1.25cm of rl1] (rl) {};
-%     \draw[dashed] (ll) -- (rl);
-
-%     \draw<1> (olcm) edge[bend left, ->,thick] node[right, near start] { } (lcm);
-%     \draw<2-> (olcm) edge[bend left, ->,thick] node[right, near start] {$id$} (lcm);
-%   \end{tikzpicture}
-% \end{column}
-% \end{columns}
-% \end{frame}
-
-% \begin{frame}
-%   \frametitle{Evaluating the definition}
-%   \begin{enumerate}
-%   \item {\itshape Intuitive}\\
-%     \onslide<2->{Yes, call stack corresponds exactly to our intuition.}
-% \item {\itshape Useful for reasoning}\\% we should be able to use WBCF and LSE
-%                                 % when reasoning about correctness and security
-%                                 % of programs using \stktokens{}.
-%   \onslide<3->{Yes, ...}
-% \item {\itshape Reusable in secure compiler chains}\\% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
-%   \onslide<4->{Yes, fully-abstract compilations compose vertically, so oLCM can
-%     be used as a target for other compilation phases.}
-% \item {\itshape Arguably "complete"}\\% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
-%   \onslide<5->{Yes, ...}
-% %\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
-% \end{enumerate}
-% \end{frame}
-
-% \begin{frame}
-%   \frametitle{The \stktokens{} calling convention}
-%   \begin{onslide}<2->
-%     ... is explained in the paper.
-%   \end{onslide}
-% \end{frame}
+\begin{frame}
+  \frametitle{Proof sketch for a realistic secure compiler}
+  \begin{center}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5},distance=1cm]
+    % \draw<1-6> (-5,0) node {};
+    % \draw (5,0) node {};
+    % \draw (0,-14) node {};
+    \draw<1-> node[draw=black,rounded corners,above] (source) {High-level language};
+    \draw<1-> node[draw=black,rounded corners,below=of source] (mca) {Machine code w/ many abstractions};
+    \draw<1-> node[draw=black,rounded corners,below=1cm of mca] (mcir) {...};
+    \draw<1-> node[draw=black,rounded corners,below=1cm of mcir] (mcs) {Machine code w/ built-in stack};
+    \draw<1-> node[draw=black,rounded corners,below=of mcs] (target) {Machine code};
+
+    \draw<1-> (source) edge[bend left, ->] (mca);
+    \draw<1-> (mca) edge[bend left, ->] (mcir);
+    \draw<1-> (mcir) edge[bend left, ->] (mcs);
+    \draw<1-> (mcs) edge[bend left, ->] (target);
+    \begin{visibleenv}<2->
+      \draw node[rectangle, fit=(mca) (target), draw=myred,inner sep=10pt, rounded corners] (lowlevel) {};
+      \draw (lowlevel) node[rectangle, inner sep=10pt, opaque=0.5, fill=myred,fill opacity=0.75, text opacity=1,rotate=-30] (text) {Overlay semantics stack};
+    \end{visibleenv}
+  \end{tikzpicture}                      
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Full-abstraction proof sketch}
+  \centering
+    \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
+    % \draw[help lines,yellow] (0,0) grid (10,7);
+    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
+
+    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
+    \node at (5,4) { $\mathrel{\Rightarrow}$ };
+    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
+
+    \node at (4.35,2.8) {  };
+    \node at (5,3.6) {  };
+    \node at (5.65,2.8) {  };
+
+    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
+
+    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
+
+    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
+    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
+      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
+    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
+    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
+    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
+
+    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
+    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
+    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
+
+    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
+
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em, below=0.5cm]{\small Contextual equivalence preservation} (0,1);
+  \end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Desired properties of LSE and WBCF definition}
+  \begin{enumerate}
+\item {\itshape intuitive} \label{def-prop:intuitive}
+\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
+% "Useful for reasoning" is intended to contrast our approaches in
+% ESOP18 (logical relation) and POPL19 (fully abstract overlay
+% semantics) against Abadi's use of control flow graphs.
+%   The idea is that if you have a concrete program whose correctness
+% depends on WBCF and LSE, then one should be able to rely on the
+% formalization of WBCF and LSE of the calling convention to prove
+% correctness of the program.
+%   This intuitive desirable property is satisfied for our ESOP18 paper
+% (you can use the LR to reason about your program) and our POPL19 paper
+% (you can reason in oLCM where WBCF and LSE hold, using a separate LR
+% or other tool if needed) but not for Abadi's property about "control
+% flow stays withing the static control flow graph" (how would you use
+% this for reasoning?, time-insensitive).
+\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
+\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
+\label{def-prop:complete}
+% "Arguably complete" is intended to contrast our POPL19 approach
+% against the ones in ESOP18 and Abadi's CFG approach.
+%   It is intended to mean that the formalization of WBCF and LSE should
+% arguably apply to "all properties that follow from WBCF and LSE".
+%   For POPL19, this is satisfied because it applies to any property
+% that holds in oLCM.
+%   For Abadi's CFGs, this is not satisfied, for one reason because it
+% is not timing-sensitive.
+%   For ESOP18, I don't see any reason to think that any result provable
+% using WBCF and LSE should also be provable using our LR, even though
+% there is also no reason to think it isn't.
+\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
+\end{enumerate}
+\end{frame}
 
@@
 % \end{lstlisting}
-
-\appendix
-\begin{frame}
-  \frametitle{Full-abstraction proof sketch}
-  \centering
-    \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
-    % \draw[help lines,yellow] (0,0) grid (10,7);
-    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
-
-    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
-    \node at (5,4) { $\mathrel{\Rightarrow}$ };
-    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
-
-    \node at (4.35,2.8) {  };
-    \node at (5,3.6) {  };
-    \node at (5.65,2.8) {  };
-
-    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
-
-    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
-
-    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
-    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
-    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
-    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
-    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
-
-    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
-    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
-    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
-
-    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
-
-    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em, below=0.5cm]{\small Contextual equivalence preservation} (0,1);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Desired properties of LSE and WBCF definition}
-  \begin{enumerate}
-\item {\itshape intuitive} \label{def-prop:intuitive}
-\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
-\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
-\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
-\label{def-prop:complete}
-\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{FAQ}
-  \begin{itemize}
-  \item \textit{Do you support tail calls?}
-    \begin{itemize}
-    \item Yes.
-    \end{itemize}
-  \item \textit{Do you support higher-order functions?}
-    \begin{itemize}
-    \item Yes.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\end{document}p
+\end{document}

### 2019-02-25 local-capability-machine/arxiv/local-capability-machine.tex
[TOPLAS] Add arxiv tech report files

--- 2018-04-27 local-capability-machine/local-capability-machine.tex
@@
 
-\author{Lau Skorstengaard\\Aarhus University\\\texttt{lask@cs.au.dk}
+\author{Lau Skorstengaard\\Aarhus University\\\texttt{lau@cs.au.dk}
         \and 
-        Dominique Devriese\\imec-DistriNet, KU Leuven\\\texttt{dominique.devriese@cs.kuleuven.be}
+        Dominique Devriese\\Vrije~Universiteit~Brussel\\\texttt{dominique.devriese@vub.be}
         \and 
@@
               \item[\texttt{fetch} $r$ $f$] load the entry of the linking table corresponding to $f$ to register $r$.\\
-One possible malloc implementation (r\_t1 and r\_t2 are registers in RegName\_t).
+One possible fetch implementation (r\_t1 and r\_t2 are registers in RegName\_t).
 \begin{lstlisting}
@@
       \begin{enumproof}
-        \item $\npair{\reg_0\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}}(c_\stk)} \in \stdvr(W)$ \\
+        \item $\npair{c_\stk} \in \stdvr(W)$ \\
           The assumptions on $c_\stk$ and $\ms_\stk$ in the lemma entail
@@
   \item $\memSat{\ms}{W}$
-  \item $\npair{\stdcap} \in stdvr(W)$
+  \item $\npair{\stdcap} \in \stdvr(W)$
   \item $\start \leq \addrend$
@@
   \label{lem:malloc-in-vr}
-  If $W(r) \futurestr \iota_{\malloc,0}$, then
+    For $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
   $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
@@
     (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus \ms_f)$
-    and $\Phi \step[i] (\halted,\Phi')$,\\
+    and $\Phi \step[i] (\halted,\mem')$,\\
     Suffices: $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-    $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+    $\mem' = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
     By: definition of $\observations(W)$ \label{suff-after-eval}
@@
     $(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\hs \uplus
-    \ms_f)\neq (\halted,\Phi')$ for any $\Phi'$. \\
+    \ms_f)\neq (\halted,\mem')$ for any $\mem'$. \\
     Therefore, assume w.l.o.g. that $i = 1+i'$,
     \begin{equation*}
-      \Phi \step \var{conf'} \step[i'] (\halted,\Phi')
+      \Phi \step \var{conf'} \step[i'] (\halted,\mem')
     \end{equation*} \label{step:ip-non-zero}
@@
   \item Case $\var{conf'} = \failed$: contradiction, since it is not possible
-    that $\failed \step[i'] (\halted,\Phi')$. 
-  \item Case $\var{conf'} = (\halted,\Phi)$:
-    \begin{enumproof}
-    \item Then $i' = 0$ and $\Phi' =
-      \Phi$\\
-      Follows from $(\halted, \Phi) \step[i'] (\halted,\Phi)$
+    that $\failed \step[i'] (\halted,\mem')$. 
+  \item Case $\var{conf'} = (\halted,\mem)$:
+    \begin{enumproof}
+    \item Then $i' = 0$ and $\mem' =
+      \mem$\\
+      Follows from $(\halted, \mem) \step[i'] (\halted,\mem)$
     \item For $W' = W$, $\ms_r = \emptyset$ and $\ms' = \ms$, we have that
-      $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and
+      $\mem = \hs' \uplus \hs_r \uplus \ms_f$ and
       $\heapSat[\hs']{n'-1}{W'}$ (using Lemma~\ref{lem:heap-sat-dc}).
@@
     \item $\var{newPc} = ((\perm',\gl'),\start',\addrend',\addr' + 1)$ and
-      $\memreg[\Phi''](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$
+      $\memreg[\Phi](\pcreg) = ((\perm',\gl'),\start',\addrend',\addr')$
     \end{itemize}
@@
     \item $\exists W' \futurestr W$, $\hs_r$, $\hs' \ldotp$
-      $\memheap[\Phi'] = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
+      $\mem = \hs' \uplus \hs_r \uplus \ms_f$ and $\heapSat[\hs']{n'-i}{W'}$\\
       By: definition of $\observations(W)$ using the previous step and the
-      evaluation $\var{conf'}\step[i'] (\halted,\Phi')$ from
+      evaluation $\var{conf'}\step[i'] (\halted,\mem')$ from
       Step~\ref{step:ip-non-zero}.
@@
 
-\subsubsection{Helper lemmas about the stack}
+\subsubsection{Stack helper lemmas}
 \begin{lemma}

### 2019-03-01 toplas/paper.tex
[TOPLAS] Make small textual improvements

@@
 \newcommand{\Rels}{\plaindom{Rels}}
-\newcommand{\States}{\plaindom{WState}}
+\newcommand{\States}{\plaindom{RState}}
 \newcommand{\RegionNames}{\plaindom{RegionName}}
@@
 To avoid tedious details irrelevant to the focus of this paper, we assume an
-infinite address space and unbounded integers (see
-\sectionname~\ref{sec:discussion} for more on this).
+infinite address space and unbounded integers (see \sectionname~\ref{sec:discussion} for a discussion of these assumptions).
 
@@
     \hline
-    \itoplas{$\subseg{\refreg{r_1}}{r_2}{r_3}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{w}}}$} &  \itoplas{$\memreg(r_1) = \stdcap$ and for $i \in \{2,3\}$ $n_i = \memreg(\rv_i)$ and $n_2 \in \nats$ and $\start \leq n_2$ and $n_3 \leq \addrend$ where either $n_3 \in \nats$ or ($n_3=\inftyend$ and $\addrend = \infty$) and $\perm \neq \entry$ and $w = ((\perm,\gl),n_1,n_2,\addr)$}\\
+    \itoplas{$\subseg{\refreg{r_1}}{r_2}{r_3}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{w}}}$} &  \itoplas{$\memreg(r_1) = \stdcap$ and for $i \in \{2,3\}$ $n_i = \memreg(\rv_i)$ and $n_2 \in \nats$ and $\start \leq n_2$ and $n_3 \leq \addrend$ where either $n_3 \in \nats$ or ($n_3=\inftyend$ and $\addrend = \infty$) and $\perm \neq \entry$ and $w = ((\perm,\gl),n_2,n_3,\addr)$}\\
     \hline
@@
     \end{tikzpicture}
-    \caption{The stack just before we return to the adversary. Our stack frame
+    \caption{Just before we return to the adversary. Our stack frame
       is the topmost, and we do not have access to the adversary stack frame.}
@@
     \end{tikzpicture}
-    \caption{The stack just after we returned to the adversary but did not take
+    \caption{Just after we returned to the adversary but did not take
       care to clear our local stack frame.}
@@
     \end{tikzpicture}
-    \caption{The stack just after we returned to the adversary, and we did take
+    \caption{Just after we returned to the adversary, and we did take
       care to clear our local stack frame.}
@@
     \end{tikzpicture}
-  \caption{The stack after an adversary has called us, and we have called the adversary.}
+  \caption{After an adversary has called us, and we have called the adversary.}
   \label{fig:illu-call-adv}
@@
     \end{tikzpicture}
-  \caption{The stack after the adversary has returned to us, and we have
+  \caption{After the adversary has returned to us, and we have
     returned from the first adversary call. At this point, the adversary has
@@
     \end{tikzpicture}
-  \caption{The stack after the adversary has returned to us, and we have
+  \caption{After the adversary has returned to us, and we have
     returned from the first adversary call after we took care to clear the
@@
 In our correctness proofs, worlds allow us to rely on the memory having a particular shape, but sometimes we will also have to guarantee that the memory has a certain shape.
-Essentially, a world is a collection of invariants. The memory satisfies the world when part of the memory satisfy each of the invariants.
+Essentially, a world is a collection of invariants.
+The memory satisfies the world when part of the memory satisfy each of the invariants.
 
@@
 
-For technical reasons, we do not actually remove a revoked temporary region from the world, but we turn it into a special revoked region that exists for this purpose.
+We need the future world relation to be extensional, so we do not actually remove a revoked temporary region from the world, but we turn it into a special revoked region that exists for this purpose.
 Such a revoked region contains no state transition system and puts no requirements on the memory.
@@
   In order to be able to solve the recursive domain equation, we want the future world relations to be extensional.
-  This means a future world cannot forget about region names in a past world.
+  Intuitively, the future world relation specifies how memory protocols may change over time.
+  If the future world relation was not extensional, then the protocols could just disappear entirely regardless of what they promise.
+  A piece of memory can still be repurposed under the right circumstances, i.e., if the protocol allows it under the future world relation used.
+  In practice, the extensionality of future worlds means that a future world cannot forget about region names in a past world.
   This is the reason we use masks rather than deleting the regions altogether.
@@
 %   respect them itself too.} 
-The logical relation defines semantically when values, program
-counters, and configurations are capability safe. The logical relation is defined
-found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and
-we provide some explanations in the following paragraphs. For space
-reasons, we omit some definitions and explain them only verbally, but
-precise definitions can be found in the appendix.
+The logical relation defines semantically when values, program counters, and configurations are capability safe.
+The logical relation is defined found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and we provide some explanations in the following paragraphs.
+For space reasons, we omit some definitions and explain them only verbally, but precise definitions can be found in the appendix.
+The logical relation is recursively defined, so we encourage first time readers to just read the section in its entirety and do a second read afterwards.
 

### 2019-03-06 toplas/paper.tex
Add an ML-version of the awkward example and use it to explain the assembly version.

@@
 The final example, \texttt{g1} with \texttt{f4}, is a faithful translation of a tricky example known from the literature (known as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}.
-It consists of two parts, \texttt{g1} and \texttt{f4}.
-The program \texttt{g1} is a closure generator that generates closures with one variable $x$ set to $0$ in its environment and \texttt{f4} as the program (note we omit some calling convention security measures because the stack is not used in the closure generator).
-The program\texttt{f4} expects one argument, a callback.
-\texttt{f4} sets $x$ to $0$ and calls the callback.
-When the callback returns, \texttt{f4} sets $x$ to $1$ and calls the callback again.
-When the callback returns again, \texttt{f4} asserts $x$ is $1$ and returns.
+For comparison, we show a version of the original example in Figure~\ref{fig:awkward}, highlighting the code locations corresponding to \texttt{g1} and \texttt{f4}.
+
+\begin{wrapfigure}{R}{.35\textwidth}
+  \centering
+  \begin{tabular}{c} % see https://tex.stackexchange.com/questions/5818/how-to-center-a-listing
+  \begin{lstlisting}[gobble=2,escapeinside=||]
+  fun _ => 
+  |\color{gray}{g1:}| let x = ref 0 in
+      fun adv =>
+  |\color{gray}{f4:}|   x := 0;
+        adv();
+        x := 1;
+        adv();
+        assert(x == 1)
+  \end{lstlisting}
+  \end{tabular}
+  \caption{The original awkward example from \citet{pitts_operational_1998,Dreyer:jfp12}, in ML notation.}
+  \label{fig:awkward}
+\end{wrapfigure}
+
+Let us first look at this ML program.
+At the top-level, it is a lambda function that can be invoked by the context.
+When it is invoked, it allocates a fresh mutable variable $x$ of integer type, initially containing the value $0$.
+Next, the function returns a second closure (let's call this $\textit{cl}_x$) that can be invoked by the context whenever it chooses.
+$\textit{cl}_x$ itself takes a callback function \texttt{adv} that it will invoke twice, after setting $x$ to $0$ and $1$ respectively.
+After the second invocation of \texttt{adv}, $\textit{cl}_x$ will verify that $x$ is still set to $1$.
+
+This last assertion, at line 8 of Figure~\ref{fig:awkward}, should never fail.
+From the code, this seems natural, since \texttt{adv} does not have access to $x$.
+Therefore, after the second invocation at line 7, $x$ should still be in the state it was before that invocation.
+However, consider that \texttt{adv} may have access to $\textit{cl}_x$ and within that second invocation of \texttt{adv}, it may invoke $\textit{cl}_x}$ again.
+Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
+If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up at line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
+Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
+In other words, if we are able to prove that the assertion will never fail, this attack, and other similar attacks, on well-bracketed control flow are adequately prevented.
+
+Our low-level version of the awkward example in Figure~\ref{fig:prog-f3-and-g1}, consists of two parts, \texttt{g1} and \texttt{f4}, corresponding to the code locations marked in Figure~\ref{fig:awkward}.
+The program \texttt{g1} corresponds to the top-level closure in the ML code.
+It is a closure generator that generates closures with a mutable variable $x$ set to $0$ in its environment and \texttt{f4} as the program (note that we omit some calling convention security measures because the stack is not used in the closure generator).
+
+The program \texttt{f4} expects one argument, the callback \texttt{adv}, in \texttt{r1}.
+\texttt{f4} sets $x$ to $0$ and invokes \texttt{adv}.
+When it returns, \texttt{f4} sets $x$ to $1$ and calls \texttt{adv} again.
+When it returns the second time, \texttt{f4} asserts $x$ is still $1$ and returns.
+
 This example is more complicated than the previous ones because it involves a closure invoked by the adversary and an adversary callback invoked by us.
-\itoplassug{ Suggestion: Maybe give some more intuition as to why it is difficult.
-} As explained in \sectionname~\ref{sec:stack-and-return-pointer}, we need to check that (1) the stack pointer that the closure receives from the adversary has write-local permission and (2) the adversary callback is global.
-
-To illustrate how subtle this program is, consider how an adversary could try to
-make the assertion fail. In the second callback an adversary can get to the
-first callback by invoking the closure one more time. If there were any way for
-the adversary to transfer the return pointer from the point where it reinvokes
-the closure to where the closure reinvokes the callback, then the assertion
-could be made to fail. Similarly, if there were any way for the adversary to
-store a stack pointer or trick the trusted code into preserving it across an
-invocation, the assertion could be made to fail too. However, our calling
-convention prevents any of this from happening, as we prove in the following
-lemma.
+As explained in \sectionname~\ref{sec:stack-and-return-pointer}, we need to check that (1) the stack pointer that the closure receives from the adversary has write-local permission and (2) the adversary callback is global.
+
+The attack that we explained above is actually more natural at this low-level.
+The callback \texttt{adv} now gets a return capability.
+As explained, it could invoke $\mathit{cl}_x$ again, during its second invocation, with a second callback function $\texttt{adv}'$ for $\mathit{cl}_x$ to invoke.
+If \texttt{adv} had some way of passing its return capability to $\texttt{adv}'$ (by storing it somewhere on the stack or in the heap, or hide it an unused register), then the assertion could be made to fail.
+However, our calling convention prevents any of this from happening, as we prove in the following lemma.
 

### 2019-03-06 toplas/paper.tex
minor

@@
   \begin{itemize}
-  \item Proof sketches for the \nameref{thm:ftlr} and Lemma~\ref{lem:correctness-g1}, the correctness lemma for the awkward example.
-  \item Figures that illustrate central parts of the calling convention.
-  \item A section about malloc.
+  \item We have added proof sketches for the \nameref{thm:ftlr} and for the correctness lemma for the awkward example
+    (Lemma~\ref{lem:correctness-g1}).
+  \item We have added figures that illustrate central parts of the calling convention.
+  \item A section about malloc has been added.
     Specifically, we provide the specification for the malloc used in the examples of the paper.
-  \item A section about macro instructions with descriptions of all the macros and the implementation of \texttt{scall}.
-  \item A section on reasoning about programs that run on a capability machine.
+  \item A section about macro instructions with descriptions of all the macros and the implementation of \texttt{scall} has been added.
+  \item We have added a section on reasoning about programs that run on a capability machine.
     This section explains how one reasons about common scenarios that arise in programs on a capability machine, and, in particular, how the logical relation is used.
     It also introduces a number of lemmas that prove recurring bits once and for all.
-  \item Details previously found only in the technical appendix~\citep{technical_appendix} has been moved to the paper.
-    Further, many of the details omitted in the paper has been added to the paper's appendix.
+  \item We have expanded on the explanation of the awkward example. 
+  \item Details previously found only in the technical appendix~\citep{technical_appendix} have been moved to the paper.
 \end{itemize}

### 2019-03-06 toplas/paper.tex
[TOPLAS] Fix compilation issue

@@
 Therefore, after the second invocation at line 7, $x$ should still be in the state it was before that invocation.
-However, consider that \texttt{adv} may have access to $\textit{cl}_x$ and within that second invocation of \texttt{adv}, it may invoke $\textit{cl}_x}$ again.
+However, consider that \texttt{adv} may have access to $\textit{cl}_x$ and within that second invocation of \texttt{adv}, it may invoke $\textit{cl}_x$ again.
 Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.

### 2019-03-06 toplas/paper.tex
[TOPLAS] Add keywords to ML awkward example

@@
   \begin{tabular}{c} % see https://tex.stackexchange.com/questions/5818/how-to-center-a-listing
-  \begin{lstlisting}[gobble=2,escapeinside=||]
+  \begin{lstlisting}[gobble=2,escapeinside=||,keywords={fun, let, in, ref, assert}]
   fun _ => 
@@
 Next, the function returns a second closure (let's call this $\textit{cl}_x$) that can be invoked by the context whenever it chooses.
-$\textit{cl}_x$ itself takes a callback function \texttt{adv} that it will invoke twice, after setting $x$ to $0$ and $1$ respectively.
+$\textit{cl}_x$ itself takes a callback function \texttt{adv} that it will invoke twice after setting $x$ to $0$ and $1$ respectively.
 After the second invocation of \texttt{adv}, $\textit{cl}_x$ will verify that $x$ is still set to $1$.

### 2019-03-06 toplas/paper.tex
[TOPLAS] Make minor changes to ML text

@@
 At the top-level, it is a lambda function that can be invoked by the context.
-When it is invoked, it allocates a fresh mutable variable $x$ of integer type, initially containing the value $0$.
+When it is invoked, it allocates a fresh mutable variable $x$ of integer type that initially contains the value $0$.
 Next, the function returns a second closure (let's call this $\textit{cl}_x$) that can be invoked by the context whenever it chooses.
@@
 
-This last assertion, at line 8 of Figure~\ref{fig:awkward}, should never fail.
-From the code, this seems natural, since \texttt{adv} does not have access to $x$.
-Therefore, after the second invocation at line 7, $x$ should still be in the state it was before that invocation.
-However, consider that \texttt{adv} may have access to $\textit{cl}_x$ and within that second invocation of \texttt{adv}, it may invoke $\textit{cl}_x$ again.
+The assertion on line 8 of Figure~\ref{fig:awkward} should never fail.
+From the code, this seems natural since \texttt{adv} does not have access to $x$.
+Therefore after the second invocation on line 7, $x$ should still be in the state it was before that invocation.
+However, \texttt{adv} could have access to $\textit{cl}_x$ and could invoke $\textit{cl}_x$ again within the second invocation of \texttt{adv}.
 Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
-If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up at line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
+If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
 Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
@@
 As explained, it could invoke $\mathit{cl}_x$ again, during its second invocation, with a second callback function $\texttt{adv}'$ for $\mathit{cl}_x$ to invoke.
-If \texttt{adv} had some way of passing its return capability to $\texttt{adv}'$ (by storing it somewhere on the stack or in the heap, or hide it an unused register), then the assertion could be made to fail.
-However, our calling convention prevents any of this from happening, as we prove in the following lemma.
+If \texttt{adv} had some way of passing its return capability to $\texttt{adv}'$ (by storing it on the stack or in the heap, or hide it in an unused register), then the assertion could be made to fail.
+However, our calling convention prevents any of this from happening as we prove in the following lemma.
 

### 2019-03-06 toplas/paper.tex
[TOPLAS] Tone down simplification text

@@
 It is inspired by both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and CHERI~\citep{Watson2015Cheri}.
-To avoid tedious details irrelevant to the focus of this paper, we assume an
+For simplicity, we assume an
 infinite address space and unbounded integers (see \sectionname~\ref{sec:discussion} for a discussion of these assumptions).

### 2019-03-08 presentations/cambridge-18-03-2019/presentation.tex
[Presentation] First commit, Cambridge '19 presentation

--- 2019-01-24 presentations/popl19-01-2019/presentation.tex
@@
 
+\tikzset{capa/.style={decoration={markings,mark=at position 0 with %
+      {\fill[black] circle(1.5pt);}}}}
+
 % \draw <coord> pic {lock};
@@
 
+\begin{frame}[fragile]
+  \frametitle{Well-bracketed control flow and local state encapsulation}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+\begin{lstlisting}[language=C, escapechar=!]
+!\tikz[remember picture] \node[] (a) {};!void a()       !\tikz[remember picture] \node[] (bt) {};! 
+ {
+   !\tikz[remember picture] \node[] (pc6) {};!...
+   !\tikz[remember picture] \node[] (pc7) {};!return;!\tikz[remember picture] \node[] (ret) {};!      
+ }              !\tikz[remember picture] \node[] (bb) {};!
+
+ void b()
+ {
+   !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
+   !\tikz[remember picture] \node[] (pc2) {};!a();
+   !\tikz[remember picture] \node[] (pc3) {};!...
+   !\tikz[remember picture] \node[] (pc4) {};!a();
+   !\tikz[remember picture] \node[] (pc5) {};!return;   
+ }
+\end{lstlisting}
+% Mention that our work support high-order
+% No good definition. We appeal to you intuitive definition
+      \begin{tikzpicture}[remember picture, overlay]
+%        \draw<1> node { };
+        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=0.2cm] (brace) [midway] {};
+        \draw<2> node[right = -0cm of brace, text width = 3cm] {Function \texttt{a} cannot access variable \texttt{x}};
+
+        \draw<3> ($(pc4) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc4);
+        \draw<4> ($(a) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (a);
+        \draw<5> ($(pc6) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc6);
+        \draw<6> ($(pc7) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc7);
+        \draw<7-> node[below right = of ret] (w) {\huge ?};
+        \draw<7-> (ret) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202},bend left] (w);
+        \draw<8> ($(pc5) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc5);
+        \draw<9> ($(pc3) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc3);
+      \end{tikzpicture}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \only<2>{Local-state encapsulation (LSE)}
+      \only<3->{Well-bracketed control flow (WBCF)}
+    \end{column}
+    % What is LSE and WBCF
+    % Quick example?
+  \end{columns}
+\end{frame}
+
 % Secure compilations: If security properties in target can be proven in source.
@@
     % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);
+    \draw (20,14) node {\phantom{heap memory}};
 
@@
     \stdstackstart
+    \draw (20,14) node {\phantom{heap memory}};
+    \regfile
 
@@
     % prepare call: store old return capability
-    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\tiny old return cap code} (6,6.5);
-    \begin{onlyenv}<+-+(2)>
+    \begin{onlyenv}<+-+(3)>
       \draw[fill=gray!50] (0,5.5) rectangle node {\tiny old return cap data} (6,6);
@@
     \end{onlyenv}
+    \draw<+-+(2)>[fill=gray!50] (0,6) rectangle node {\tiny old return cap code} (6,6.5);
     \begin{onlyenv}<+->
@@
 \begin{frame}
-  \frametitle{\stktokens{} prevent the attack}
+  \frametitle{Linear capabilities}
+  \begin{columns}
+    \begin{column}{.6\textwidth}
+      \begin{tikzpicture}[scale=.45, every node={scale=.5}]
+        \scope
+          \clip (-.1,-.1) rectangle (6.1,14.9);
+          \fill[draw,fill=white] (0,0) rectangle (6,15);
+        \endscope
+        % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward}
+        % (-2,15);
+        \foreach \X in {0,0.5,...,14} {
+          \fill[draw,fill=white] (0,\X) rectangle (6,\X+.5) node[pos=.5] { };
+        };
+        \draw (-0.5,0.25) node[align=right] { \scriptsize 0 };
+        \draw (-0.5,4.25) node[align=right] { \scriptsize 8 };
+        \draw (-0.5,8.25) node[align=right] { \scriptsize 16 };
+        \draw (-0.5,12.25) node[align=right] { \scriptsize 24 };
+
+  \begin{scope}
+    \clip (11.9,2.6) rectangle (13.6,11.6);
+    \foreach \X in {0,1.5,...,7.5} {
+      \fill[draw,fill=white] (12,10 - \X) rectangle (13.5,11.5-\X) node[pos=.5] { };
+    };
+  \end{scope}
+  \draw (12.75,12.25) node[align=center] {Register File};
+  \draw (6.5,14.325) node[anchor=west, align=left] {Memory}; \draw[->] (-1.5,0) -- (-1.5,4.5);
+
+  % Mem data
+  \draw<1-> (3,0.75) node { \scriptsize 42 };
+  \draw<1-> (3,1.25) node { \scriptsize 2 };
+  \draw<1-> (3,1.75) node { \scriptsize 14 };
+  \draw<1-> (3,2.75) node { \scriptsize 10 };
+  \draw<1-> (3,2.25) node { \scriptsize 19 };
+  \draw<1-> (3,3.75) node { \scriptsize 0 };
+  \draw<1-> (3,3.25) node { \scriptsize 0 };
+  \draw<1-> (3,4.25) node { \scriptsize 0 };
+  \draw<1-> (3,4.75) node { \scriptsize 0 };
+  \draw<1-> (3,5.75) node { \scriptsize 3 };
+  \draw<1-> (3,6.25) node { \scriptsize 20 };
+  \draw<1-> (3,6.75) node { \scriptsize 21 };
+  \draw<1-> (3,7.25) node { \scriptsize 11 };
+  \draw<1-> (3,7.75) node { \scriptsize 10 };
+  \draw<1-> (3,8.25) node { \scriptsize 15 };
+  \draw<1-> (3,8.75) node { \scriptsize 19 };
+  \draw<1-> (3,9.25) node { \scriptsize 29 };
+  \draw<1-> (3,9.75) node { \scriptsize 0 };
+  \draw<1-> (3,10.25) node { \scriptsize 0 };
+  \draw<1-> (3,10.75) node { \scriptsize 0 };
+  \draw<1-> (3,11.25) node { \scriptsize 0 };
+  \draw<1-> (3,11.75) node { \scriptsize 0 };
+  \draw<1-> (3,12.75) node { \scriptsize 0 };
+  \draw<1-> (3,13.25) node { \scriptsize 0 };
+  \draw<1-> (3,13.75) node { \scriptsize 64 };
+  \draw<1-> (3,14.25) node { \scriptsize 5 };
+  % Reg data
+  \draw<1-> (12.75,10.75) node { \scriptsize 8 };
+  \draw<1-> (12.75,7.75) node { \scriptsize 24 };
+  \draw<1-> (12.75,6.25) node { \scriptsize 16 };
+  \draw<1-> (12.75,4.75) node { \scriptsize 2 };
+
+  % Mem pointers
+  \draw<1> (3,12.25) node { \scriptsize 16 };
+  \draw<1> (3,5.25) node { \scriptsize 9 };
+  \draw<1> (3,0.25) node { \scriptsize 8 };
+  % Reg pointers
+  \draw<1> (12.75,9.25) node { \scriptsize 0 };
+
+  % Mem pointers HL
+  \draw<2> (3,12.25) node { \color{red} \scriptsize 16 };
+  \draw<2> (3,5.25) node { \color{red} \scriptsize 11 };
+  \draw<2> (3,0.25) node { \color{red} \scriptsize 8 };
+  % Reg pointers HL
+  \draw<2> (12.75,9.25) node { \color{red} \scriptsize 0 };
+
+  \draw<3> (12.75,9.25) node { };
+
+  % Mem capabilities
+  \draw<4->[capa] (3,12.25) edge[postaction={decorate}] (6.1,12.25 );
+  \draw<4> (6.1,12.25) edge[-latex,in=0,out=0] (6.1,8.25);
+  \draw<4->[capa] (3,5.25) edge[postaction={decorate}] (6.1,5.25);
+  \draw<4> (6.1,5.25) edge[-latex,in=0,out=0] (6.1,6.75);
+  \draw<4->[capa] (3,0.25) edge[postaction={decorate}] (6.1,0.25);
+  \draw<4> (6.1,0.25) edge[-latex,in=0,out=0] (6.1,4.25);
+
+  % Reg pointers HL
+  \draw<4->[capa] (12.75,9.25) edge[postaction={decorate}] (12.0,9.25);
+  \draw<4> (12.0,9.25) edge[-latex,in=0,out=180] (6.1,0.25);
+
+  % Mem capabilities \w auth
+  \draw<5> (6.1,12.25) edge[-latex,in=0,out=0] (6.5,8.25);
+  \draw<5->[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.55,mirror},yshift=0pt] (6.3,5.5) -- (6.3,10.5) node (e) [black,midway,xshift=0.8cm] {};
+
+  \draw<5> (6.1,5.25) edge[-latex,in=0,out=0] (6.3,6.75);
+  \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,6) -- (6.1,7.5) node (e) [black,midway,xshift=0.8cm] {};
+
+  \draw<5> (6.1,0.25) edge[-latex,in=0,out=0] (6.5,4.25);
+  \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.3,3) -- (6.3,5.5) node (e) [black,midway,xshift=0.8cm] {};
+
+  % Reg pointers HL \w auth
+  \draw<5> (12.0,9.25) edge[-latex,in=0,out=180] (6.3,0.355);
+  \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.1,mirror},yshift=0pt] (6.1,0) -- (6.1,3.5) node (e) [black,midway,xshift=0.8cm] {};
+
+  % Mem capabilities \w auth & perm
+  \draw<6-> (6.1,12.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{r}} (6.5,8.25);
+
+  \draw<6-> (6.1,5.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{rw}} (6.3,6.75);
+
+  \draw<6-> (6.1,0.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{r}} (6.5,4.25);
+
+  % Reg pointers HL \w auth & perm
+  \draw<6-> (12.0,9.25) edge[-latex,in=0,out=180] node[right] {\scriptsize \textsc{rx}} (6.3,0.355);
+\end{tikzpicture}
+\end{column}
+\begin{column}{0.4\textwidth}
+  \onslide<3->{Replace pointers with capabilities}
+  \begin{itemize}
+  \item<4-> Pointer
+  \item<5-> Range of authority
+  \item<6-> Permission
+  \end{itemize}
+  \onslide<7->{Dynamically enforced}
+  \begin{itemize}
+  \item<8-> Cap. aware instructions
+  \item<9-> Tagged memory
+  \end{itemize}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{\stktokens{} prevents the attack}
   % Emphazise that it is dynamically enforced.
@@
     \stdstackstart
-    \draw (17.5,5.5) node {register file};
-    \begin{scope}
-      \clip (16.9,1.1) rectangle (18.1,5.1);
-      \foreach \x in {1,2,...,4}
-      {
-        \draw[fill=white] (17,\x) rectangle (18,\x+1) node[pos=.5,color=mygreen] {};
-      };
-    \end{scope}
-    \node (r1) at (17.75,4.5) {};
-    \node (r2) at (17.75,3.6) {};
+    \draw (20,14) node {\phantom{heap memory}};
+
+    \regfile
 
@@
 % \begin{frame}
+%   \frametitle{\stktokens{} enforces single stack}
+%   \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+%     % recurrent parts
+%   \scope
+%     \clip (-.1,-.1) rectangle (6.1,15.1);
+%     \fill[fill=white] (0,0) rectangle (6,15);
+%     \draw (0,0) -- (0,15);
+%     \draw (6,0) -- (6,15);
+%     \draw[fill=gray!50] (0,-.5) rectangle (6,1.5) node[pos=.5,color=black] {\footnotesize higher stack frames...};
+%   \endscope
+%   % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
+
+%   \begin{scope}
+%     \fill[gray!50] (13,0) rectangle (18,13);
+%     \draw (13,0) -- (13,13);
+%     \draw (18,0) -- (18,13);
+%     \draw (15.5,14) node {heap};
+%   \end{scope}
+
+% \note{Caller calls callee (split stack pointer to make ret-ptr-d)}
+%   \begin{onlyenv}<+>
+%     \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+%   \end{onlyenv}
+%   \begin{onlyenv}<.>
+%     \capbrace[sp]{(6.1,1.5)}{(6.1,15)}
+%     \draw node[right=0cm of sp] { sp };
+%   \end{onlyenv}
+
+%   \begin{onlyenv}<+->
+%     \inactsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+%   \end{onlyenv}
+%   \begin{onlyenv}<.>
+%     \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+%     \draw[fill=white] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+%   \end{onlyenv}
+%   \begin{onlyenv}<.>
+%     \capbrace[sp2]{(6.1,4.5)}{(6.1,15)}
+%     \capbrace[sp]{(6.1,1.5)}{(6.1,4.5)}
+%     \draw node[right=0cm of sp2] { sp2 };
+%     \draw node[right=0cm of sp, fill=gray!50] { ret-ptr-d };
+%     \draw (13,6.5) edge[myred,bend right,thick,->,dashed] (sp);
+%   \end{onlyenv}
+
+%   \begin{onlyenv}<+>
+%     \capbrace[sp3]{(6.1,9.5)}{(6.1,15)}
+%     \capbrace[sp2]{(6.1,4.5)}{(6.1,9.5)}
+%     \capbrace[sp]{(6.1,1.5)}{(6.1,4.5)}
+%     \draw node[right=0cm of sp3] { sp3 };
+%     \draw node[right=0cm of sp2] { sp2 };
+%     \draw node[right=0cm of sp, fill=gray!50] { ret-ptr-d };
+%     \draw (13,7.5) edge[myred,bend right,thick,->,dashed] (sp2);
+%     \draw (13,6.5) edge[myred,bend right,thick,->,dashed] (sp);
+%   \end{onlyenv}
+%   \begin{onlyenv}<.->
+%     \actsf{(0,9.5)}{(6,11.5)}{callee stack frame (1)}
+%   \end{onlyenv}
+%   \begin{onlyenv}<.>
+%     \inactadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+%     \draw[fill=gray!30] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+%   \end{onlyenv}
+
+%   \begin{onlyenv}<+>
+%     \capbrace[sp4]{(6.1,11.5)}{(6.1,15)}
+%     \capbrace[sp3]{(6.1,9.5)}{(6.1,11.5)}
+%     \capbrace[sp2]{(6.1,4.5)}{(6.1,9.5)}
+%     \capbrace[sp]{(6.1,1.5)}{(6.1,4.5)}
+%     \draw node[right=0cm of sp4] { sp4 };
+%     \draw node[right=0cm of sp3, fill=gray!50] { sp3 };
+%     \draw node[right=0cm of sp2] { sp2 };
+%     \draw node[right=0cm of sp, fill=gray!50] { ret-ptr-d };
+%     \draw (13,8.5) edge[myred,bend right,thick,->,dashed] (sp3);
+%     \draw (13,7.5) edge[myred,bend right,thick,->,dashed] (sp2);
+%     \draw (13,6.5) edge[myred,bend right,thick,->,dashed] (sp);
+%   \end{onlyenv}
+%   \begin{onlyenv}<.->
+%     \inactsf{(0,9.5)}{(6,11.5)}{callee stack frame (1)}
+%     \actadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+%   \end{onlyenv}
+%   \begin{onlyenv}<.>
+%     \actadv{(0,4.5)}{(6,7.5)}{callee stack frame}
+%     \draw[fill=white] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+%   \end{onlyenv}
+
+%   \begin{onlyenv}<+->
+%     \capbrace[sp4]{(6.1,11.5)}{(6.1,15)}
+%     \capbrace[sp3]{(6.1,9.5)}{(6.1,11.5)}
+%     \capbrace[sp]{(6.1,1.5)}{(6.1,9.5)}
+%     \draw node[right=0cm of sp4, fill=gray!50] { sp4 };
+%     \draw node[right=0cm of sp3, fill=gray!50] { sp3 };
+%     \draw node[right=0cm of sp] { sp };
+%     \draw (13,8.5) edge[myred,bend right,thick,->,dashed] (sp3);
+%   \end{onlyenv}
+%   \begin{onlyenv}<.->
+%     \inactsf{(0,9.5)}{(6,11.5)}{callee stack frame (1)}
+%     \inactadv{(0,11.5)}{(6,14.5)}{callee stack frame (2)}
+%     \actsf{(0,1.5)}{(6,4.5)}{caller stack frame}
+%     \draw[fill=white] (13,6) rectangle (18,9) node[pos=.5] {\footnotesize adv mem};
+%   \end{onlyenv}
+%   \begin{onlyenv}<+->
+%     \draw[mygreen,very thick] (2.5,10) -- (3.5,9);
+%     \draw[mygreen,very thick] (3.5,10) -- (2.5,9);
+%     \draw node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center, anchor=west] () at (4,9.5) { Base check fails!};  
+%   \end{onlyenv}
+% \end{tikzpicture}
+% \end{frame}
+
+% \begin{frame}
 %   \frametitle{\stktokens{}}
@@
 
+% \begin{frame}
+%   \frametitle{\stktokens{}: ``Base address'' check}
+%   \includegraphics[width=\textwidth]{stk-fig1}
+% \end{frame}
+
 \section{Defining well-bracketed control flow and local state encapsulation by fully-abstract overlay semantics}
-\begin{frame}[fragile]
-  \frametitle{Well-bracketed control flow and local state encapsulation}
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-\begin{lstlisting}[language=C, escapechar=!]
-!\tikz[remember picture] \node[] (a) {};!void a()       !\tikz[remember picture] \node[] (bt) {};! 
- {
-   !\tikz[remember picture] \node[] (pc6) {};!...
-   !\tikz[remember picture] \node[] (pc7) {};!return;!\tikz[remember picture] \node[] (ret) {};!      
- }              !\tikz[remember picture] \node[] (bb) {};!
-
- void b()
- {
-   !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
-   !\tikz[remember picture] \node[] (pc2) {};!a();
-   !\tikz[remember picture] \node[] (pc3) {};!...
-   !\tikz[remember picture] \node[] (pc4) {};!a();
-   !\tikz[remember picture] \node[] (pc5) {};!return;   
- }
-\end{lstlisting}
-% Mention that our work support high-order
-% No good definition. We appeal to you intuitive definition
-      \begin{tikzpicture}[remember picture, overlay]
-%        \draw<1> node { };
-        \draw<2>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=0.2cm] (brace) [midway] {};
-        \draw<2> node[right = -0cm of brace, text width = 3cm] {Function \texttt{a} cannot access variable \texttt{x}};
-
-        \draw<3> ($(pc4) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc4);
-        \draw<4> ($(a) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (a);
-        \draw<5> ($(pc6) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc6);
-        \draw<6> ($(pc7) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc7);
-        \draw<7-> node[below right = of ret] (w) {\huge ?};
-        \draw<7-> (ret) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202},bend left] (w);
-        \draw<8> ($(pc5) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc5);
-        \draw<9> ($(pc3) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc3);
-      \end{tikzpicture}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \only<2>{Local-state encapsulation (LSE)}
-      \only<3->{Well-bracketed control flow (WBCF)}
-    \end{column}
-    % What is LSE and WBCF
-    % Quick example?
-  \end{columns}
-\end{frame}
 
@@
 
-% \begin{frame}
-%   \frametitle{Fully-abstract overlay semantics}
-%   \begin{columns}
-% \begin{column}{0.5\textwidth}
-%   \begin{itemize}[<+->]
-%   \item Compile from oLCM to LCM
-%   \item Compilation is the identity
-%   \item Builtin call stack abstraction in oLCM preserved after compilation
-%   \end{itemize}
-% \end{column}
-% \begin{column}{0.5\textwidth}
-%   \begin{tikzpicture}
-%     \draw (0,0) node[draw=black, rounded corners] (olcm) {oLCM};
-%     \draw node[draw=black, rounded corners,below=2.5cm of olcm] (lcm) {LCM};
-%     \draw (olcm) edge[bend left, ->,thick] (lcm);
-%     \draw node[left = 3cm of olcm] (ll1) {};
-%     \draw node[right = 3cm of olcm] (rl1) {};
-%     \draw node[below = 1.25cm of ll1] (ll) {};
-%     \draw node[below = 1.25cm of rl1] (rl) {};
-%     \draw[dashed] (ll) -- (rl);
+\begin{frame}
+  \frametitle{Proof sketch for a realistic secure compiler}
+  \begin{center}
+  \begin{tikzpicture}[scale=.5, every node={scale=.5},distance=1cm]
+    % \draw<1-6> (-5,0) node {};
+    % \draw (5,0) node {};
+    % \draw (0,-14) node {};
+    \draw<1-> node[draw=black,rounded corners,above] (source) {High-level language};
+    \draw<1-> node[draw=black,rounded corners,below=of source] (mca) {Machine code w/ many abstractions};
+    \draw<1-> node[draw=black,rounded corners,below=1cm of mca] (mcir) {...};
+    \draw<1-> node[draw=black,rounded corners,below=1cm of mcir] (mcs) {Machine code w/ built-in stack};
+    \draw<1-> node[draw=black,rounded corners,below=of mcs] (target) {Machine code};
 
-%     \draw<1> (olcm) edge[bend left, ->,thick] node[right, near start] { } (lcm);
-%     \draw<2-> (olcm) edge[bend left, ->,thick] node[right, near start] {$id$} (lcm);
-%   \end{tikzpicture}
-% \end{column}
-% \end{columns}
-% \end{frame}
-
-% \begin{frame}
-%   \frametitle{Evaluating the definition}
-%   \begin{enumerate}
-%   \item {\itshape Intuitive}\\
-%     \onslide<2->{Yes, call stack corresponds exactly to our intuition.}
-% \item {\itshape Useful for reasoning}\\% we should be able to use WBCF and LSE
-%                                 % when reasoning about correctness and security
-%                                 % of programs using \stktokens{}.
-%   \onslide<3->{Yes, ...}
-% \item {\itshape Reusable in secure compiler chains}\\% for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.
-%   \onslide<4->{Yes, fully-abstract compilations compose vertically, so oLCM can
-%     be used as a target for other compilation phases.}
-% \item {\itshape Arguably "complete"}\\% the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program. 
-%   \onslide<5->{Yes, ...}
-% %\item {\itshape potentially scalable}% although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.
-% \end{enumerate}
-% \end{frame}
-
-% \begin{frame}
-%   \frametitle{The \stktokens{} calling convention}
-%   \begin{onslide}<2->
-%     ... is explained in the paper.
-%   \end{onslide}
-% \end{frame}
+    \draw<1-> (source) edge[bend left, ->] (mca);
+    \draw<1-> (mca) edge[bend left, ->] (mcir);
+    \draw<1-> (mcir) edge[bend left, ->] (mcs);
+    \draw<1-> (mcs) edge[bend left, ->] (target);
+    \begin{visibleenv}<2->
+      \draw node[rectangle, fit=(mca) (target), draw=myred,inner sep=10pt, rounded corners] (lowlevel) {};
+      \draw (lowlevel) node[rectangle, inner sep=10pt, opaque=0.5, fill=myred,fill opacity=0.75, text opacity=1,rotate=-30] (text) {Overlay semantics stack};
+    \end{visibleenv}
+  \end{tikzpicture}                      
+  \end{center}
+\end{frame}
 
@@
 
-% \begin{lstlisting}[basicstyle=\tiny\ttfamily]
-%     move rtmp1 42
-%     store rstk rtmp1
-%     cca rstk -1
-%     geta rtmp1 rstk
-%     split rstk rretd rstk rtmp1
-%     move rtmp1 pc
-%     cca rtmp1 -20
-%     load rtmp1 rtmp1
-%     cca rtmp1 -21
-%     cseal rretd rtmp1
-%     move rretc pc
-%     cca rretc 5
-%     cseal rretc rtmp1
-%     move rtmp1 0
-%     xjmp r1 r2
-%     getb rtmp1 rstk
-%     minus 1 rtmp1 432
-%     move 2 pc
-%     cca rtmp2 5
-%     jnz rtmp2 rtmp1
-%     cca rtmp2 1
-%     jmp rtmp2
-%     fail
-%     splice rstk rdata
-%     cca rstk 1
-%     move rtmp2 0
-% \end{lstlisting}
+\appendix
 
-\appendix
 \begin{frame}
-  \frametitle{Full-abstraction proof sketch}
-  \centering
-    \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
-    % \draw[help lines,yellow] (0,0) grid (10,7);
-    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
-
-    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
-    \node at (5,4) { $\mathrel{\Rightarrow}$ };
-    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
-
-    \node at (4.35,2.8) {  };
-    \node at (5,3.6) {  };
-    \node at (5.65,2.8) {  };
-
-    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
-
-    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
-
-    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
-    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
-    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
-    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
-    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
-
-    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
-    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
-    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
-
-    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
-
-    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em, below=0.5cm]{\small Contextual equivalence preservation} (0,1);
-  \end{tikzpicture}
+  \frametitle{\stktokens{} summary}
+  \begin{itemize}
+  \item Check the base address of the stack capability before and after calls.
+  \item Make sure that local stack frames are non-empty.
+    % Not illustrated. Trusted code must claim spot on the call stack to make
+    % sure that they are returned to.
+  \item Create token and data return capability on call: split the stack
+    capability in two to get a stack capability for your local stack frame and a
+    stack capability for the unused part of the stack. The former is sealed and
+    used for the data part of the return pair.
+  \item Create code return capability on call: Seal the old program pointer.
+  \item Reasonable use of seals: Return seals are only used to seal old program
+    pointers, every return seal is only used for one call site, and they are not
+    leaked. % If return seals are used for other things, like sealing arbitrary
+            % code pointers, then they can be used as an unsealing mechanism.
+            % If the same seal is used for multiple call sites, then we lose the
+            % guarantee that we return to the correct place. One could use the
+            % wrong return capability to return "out of order"
+            % If return seals are leaked, then we have no guarantee about what
+            % they are used for.
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}
-  \frametitle{Desired properties of LSE and WBCF definition}
-  \begin{enumerate}
-\item {\itshape intuitive} \label{def-prop:intuitive}
-\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
-\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
-\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
-\label{def-prop:complete}
-\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
-\end{enumerate}
-\end{frame}
-
-\begin{frame}
-  \frametitle{FAQ}
+  \frametitle{\stktokens{} FAQ}
   \begin{itemize}
@@
 
-\end{document}p
+\end{document}

### 2019-03-15 presentations/cambridge-18-03-2019/illu-preamble.tex
[Presentation] First commit, Cambridge '19 presentation
[Presentation] Update Cambridge '19 presentation

--- 2019-01-14 presentations/popl19-01-2019/illu-preamble.tex
@@
 \usepackage{amssymb}
+
+\newcommand{\sealsymb}[1]{
+  \draw #1 node  {\footnotesize\faLock};
+}
+
+\newcommand{\sealsymbalt}[2][]{
+  \draw #2 node {\footnotesize\faLock};
+  \draw #2 node[yshift=-1.3pt,scale=0.8]  {\tiny\textcolor{white}{#1}};
+}
+
+\newcommand{\regfile}{
+      \draw (17.5,5.5) node {register file};
+    \begin{scope}
+      \clip (16.9,1.1) rectangle (18.1,5.1);
+      \foreach \x in {1,2,...,4}
+      {
+        \draw[fill=white] (17,\x) rectangle (18,\x+1) node[pos=.5,color=mygreen] {};
+      };
+    \end{scope}
+    \coordinate (r1) at (17.5,4.5) {};
+    \node[anchor=west] at (18.1,4.5) {$r_{\mathit{stk}}$};
+    \coordinate (r2) at (17.5,3.5) {};
+    \node[anchor=west] at (18.1,3.5) {$r_2$};
+    \coordinate (r3) at (17.5,2.5) {};
+    \node[anchor=west] at (18.1,2.5) {$r_3$};
+}
 
@@
 \newcommand{\capbrace}[3][sp1]{
-  \draw [decorate,decoration={brace,amplitude=8pt,mirror,raise=3pt},yshift=0pt]
+  \draw [decorate,decoration={brace,amplitude=8pt,mirror,raise=4pt},yshift=0pt]
   #2 -- #3 node[draw=black] (#1) [black,midway,xshift=0.8cm] {};
-
 }

### 2019-03-15 presentations/cambridge-18-03-2019/presentation.tex
[Presentation] Update Cambridge '19 presentation

@@
 
+\tikzset{capar/.style={decoration={markings,mark=at position 0 with %
+      {\fill[myred] circle(1.5pt);}}}}
+
+\usepackage{tikz-cd}
+\pgfdeclarelayer{bg}
+\pgfdeclarelayer{mg}
+\pgfsetlayers{bg,mg,main}
+
 % \draw <coord> pic {lock};
@@
 
-\date{POPL, January 16, 2019}
+\date{Cambridge, March, 2019}
 
@@
     \begin{onlyenv}<.-.(3)>
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt] (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
       \draw node[right=0cm of sp1] { stack capability };
+      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
     \end{onlyenv}
@@
       % (6,0) -- (6,16) node (oldframe1) [black,midway,xshift=0.5cm] {};
-      \draw (7.5,4) node[right,fill=gray!50] (rp1l) {return capability data};
+      \begin{pgfonlayer}{bg}
+        \draw<.-.(1)> (7.5,4) node[right,fill=gray!50] (rp1l) {return capability data};
+      \end{pgfonlayer}
       \draw (7.4,4.3) pic {lock};
@@
     % prepare call: store old return capability
-    \begin{onlyenv}<+-+(3)>
+    \draw<+-+(3)>[fill=gray!50] (0,6) rectangle node {\tiny old return cap code} (6,6.5);
+    \begin{onlyenv}<+-+(2)>
       \draw[fill=gray!50] (0,5.5) rectangle node {\tiny old return cap data} (6,6);
@@
     \end{onlyenv}
-    \draw<+-+(2)>[fill=gray!50] (0,6) rectangle node {\tiny old return cap code} (6,6.5);
+
     \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.4},yshift=0pt]
-      (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.4,xshift=0.8cm] {};
-      \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return capability data};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt,aspect=.35},yshift=0pt] (6,2.5) -- (6,16) node (callerframe1) [draw=black,pos=.35,xshift=0.8cm] {};
+      \begin{pgfonlayer}{bg}
+        \draw<.-> node[fill=gray!50,right=0cm of callerframe1] (rp1l) {new return capability data};
+      \end{pgfonlayer}
+      \draw[capa] (r2) edge[->,postaction={decorate},bend left] (callerframe1);
       \draw ($(rp1l.west) + (-.1,.3)$) pic {lock};
@@
     \begin{onlyenv}<+->
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt] (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
       \draw node[right=0cm of sp1] (nsp) { new stack capability };
+      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
     \end{onlyenv}
@@
     \stdstackstart
-    \draw (20,14) node {\phantom{heap memory}};
     \regfile
@@
       (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+
       \draw node[right=0cm of sp1] { stack pointer };
+%      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
     \end{onlyenv}
@@
       \draw node[right=0cm of sp1] { stack pointer };
+      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
       \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
@@
       \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
-      \draw (18,8.25) edge[->,color=myred,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
-      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt] (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
+      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
       \draw node[right=0cm of sp1] { stack pointer };
+      \draw (18,8.25) edge[->,color=myred,very thick,out=235,in=0,near end] node[sloped, below] {read/write} (6,8);
       % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
@@
         \scope
-          \clip (-.1,-.1) rectangle (6.1,14.9);
+          \clip (-.1,.1) rectangle (6.1,14.9);
           \fill[draw,fill=white] (0,0) rectangle (6,15);
@@
         % (-2,15);
-        \foreach \X in {0,0.5,...,14} {
+        \foreach \X in {0.5,1,...,14} {
           \fill[draw,fill=white] (0,\X) rectangle (6,\X+.5) node[pos=.5] { };
         };
-        \draw (-0.5,0.25) node[align=right] { \scriptsize 0 };
-        \draw (-0.5,4.25) node[align=right] { \scriptsize 8 };
-        \draw (-0.5,8.25) node[align=right] { \scriptsize 16 };
-        \draw (-0.5,12.25) node[align=right] { \scriptsize 24 };
 
@@
           };
+
         \end{scope}
+        \coordinate (r1) at (12.75,10.75) {};
+        \node[anchor=west] at (13.6,10.75) {$r_1$};
+        \coordinate (r2) at (12.75,9.25) {};
+        \node[anchor=west] at (13.6,9.25) {$r_2$};
+        \coordinate (r3) at (12.75,7.75) {};
+        \node[anchor=west] at (13.6,7.75) {$r_3$};
+        \coordinate (r4) at (12.75,6.25) {};
+        \node[anchor=west] at (13.6,6.25) {$r_4$};
+
         \draw (12.75,12.25) node[align=center] {Register File};
-  \draw (6.5,14.325) node[anchor=west, align=left] {Memory}; \draw[->] (-1.5,0) -- (-1.5,4.5);
-
-  % Mem data
-  \draw<1-> (3,0.75) node { \scriptsize 42 };
-  \draw<1-> (3,1.25) node { \scriptsize 2 };
-  \draw<1-> (3,1.75) node { \scriptsize 14 };
-  \draw<1-> (3,2.75) node { \scriptsize 10 };
-  \draw<1-> (3,2.25) node { \scriptsize 19 };
-  \draw<1-> (3,3.75) node { \scriptsize 0 };
-  \draw<1-> (3,3.25) node { \scriptsize 0 };
-  \draw<1-> (3,4.25) node { \scriptsize 0 };
-  \draw<1-> (3,4.75) node { \scriptsize 0 };
-  \draw<1-> (3,5.75) node { \scriptsize 3 };
-  \draw<1-> (3,6.25) node { \scriptsize 20 };
-  \draw<1-> (3,6.75) node { \scriptsize 21 };
-  \draw<1-> (3,7.25) node { \scriptsize 11 };
-  \draw<1-> (3,7.75) node { \scriptsize 10 };
-  \draw<1-> (3,8.25) node { \scriptsize 15 };
-  \draw<1-> (3,8.75) node { \scriptsize 19 };
-  \draw<1-> (3,9.25) node { \scriptsize 29 };
-  \draw<1-> (3,9.75) node { \scriptsize 0 };
-  \draw<1-> (3,10.25) node { \scriptsize 0 };
-  \draw<1-> (3,10.75) node { \scriptsize 0 };
-  \draw<1-> (3,11.25) node { \scriptsize 0 };
-  \draw<1-> (3,11.75) node { \scriptsize 0 };
-  \draw<1-> (3,12.75) node { \scriptsize 0 };
-  \draw<1-> (3,13.25) node { \scriptsize 0 };
-  \draw<1-> (3,13.75) node { \scriptsize 64 };
-  \draw<1-> (3,14.25) node { \scriptsize 5 };
-  % Reg data
-  \draw<1-> (12.75,10.75) node { \scriptsize 8 };
-  \draw<1-> (12.75,7.75) node { \scriptsize 24 };
-  \draw<1-> (12.75,6.25) node { \scriptsize 16 };
-  \draw<1-> (12.75,4.75) node { \scriptsize 2 };
-
-  % Mem pointers
-  \draw<1> (3,12.25) node { \scriptsize 16 };
-  \draw<1> (3,5.25) node { \scriptsize 9 };
-  \draw<1> (3,0.25) node { \scriptsize 8 };
-  % Reg pointers
-  \draw<1> (12.75,9.25) node { \scriptsize 0 };
-
-  % Mem pointers HL
-  \draw<2> (3,12.25) node { \color{red} \scriptsize 16 };
-  \draw<2> (3,5.25) node { \color{red} \scriptsize 11 };
-  \draw<2> (3,0.25) node { \color{red} \scriptsize 8 };
-  % Reg pointers HL
-  \draw<2> (12.75,9.25) node { \color{red} \scriptsize 0 };
-
-  \draw<3> (12.75,9.25) node { };
-
-  % Mem capabilities
-  \draw<4->[capa] (3,12.25) edge[postaction={decorate}] (6.1,12.25 );
-  \draw<4> (6.1,12.25) edge[-latex,in=0,out=0] (6.1,8.25);
-  \draw<4->[capa] (3,5.25) edge[postaction={decorate}] (6.1,5.25);
-  \draw<4> (6.1,5.25) edge[-latex,in=0,out=0] (6.1,6.75);
-  \draw<4->[capa] (3,0.25) edge[postaction={decorate}] (6.1,0.25);
-  \draw<4> (6.1,0.25) edge[-latex,in=0,out=0] (6.1,4.25);
-
-  % Reg pointers HL
-  \draw<4->[capa] (12.75,9.25) edge[postaction={decorate}] (12.0,9.25);
-  \draw<4> (12.0,9.25) edge[-latex,in=0,out=180] (6.1,0.25);
-
-  % Mem capabilities \w auth
-  \draw<5> (6.1,12.25) edge[-latex,in=0,out=0] (6.5,8.25);
-  \draw<5->[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.55,mirror},yshift=0pt] (6.3,5.5) -- (6.3,10.5) node (e) [black,midway,xshift=0.8cm] {};
-
-  \draw<5> (6.1,5.25) edge[-latex,in=0,out=0] (6.3,6.75);
-  \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,6) -- (6.1,7.5) node (e) [black,midway,xshift=0.8cm] {};
-
-  \draw<5> (6.1,0.25) edge[-latex,in=0,out=0] (6.5,4.25);
-  \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.3,3) -- (6.3,5.5) node (e) [black,midway,xshift=0.8cm] {};
-
-  % Reg pointers HL \w auth
-  \draw<5> (12.0,9.25) edge[-latex,in=0,out=180] (6.3,0.355);
-  \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.1,mirror},yshift=0pt] (6.1,0) -- (6.1,3.5) node (e) [black,midway,xshift=0.8cm] {};
-
-  % Mem capabilities \w auth & perm
-  \draw<6-> (6.1,12.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{r}} (6.5,8.25);
-
-  \draw<6-> (6.1,5.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{rw}} (6.3,6.75);
-
-  \draw<6-> (6.1,0.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{r}} (6.5,4.25);
-
-  % Reg pointers HL \w auth & perm
-  \draw<6-> (12.0,9.25) edge[-latex,in=0,out=180] node[right] {\scriptsize \textsc{rx}} (6.3,0.355);
+        \draw (6.5,14.325) node[anchor=west, align=left] {Memory};
+
+        % move
+        \draw<1-2>[capa] (r2) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,11.25);
+        \draw<1-5>[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,10) -- (6.1,12.5) node (e) [black,midway,xshift=0.8cm] {};
+        \draw<3-4>[capa] (r1) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,11.25);
+        \node<3> at (r2) {0};
+
+        % store 
+        \draw<4-5>[capa] (r4) edge[->,postaction={decorate},out=180,in=0] node[above] {\tiny\textsc{rw}} (6.3,7.75);
+        \draw<4-5>[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,6.5) -- (6.1,9) node (e) [black,midway,xshift=0.8cm] {};
+        \draw<5>[capa] (3,8.25) edge[postaction={decorate},dashed] (6.3,8.25);
+        \draw<5> (6.3,8.25) edge[->,out=0,in=0,dashed] (6.3,11.25);
+        \node<5> at (r1) {0};
+
+        % split
+        \draw<6>[capa] (r3) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,10);
+        \draw<6>[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,7.5) -- (6.1,12.5) node (e) [black,midway,xshift=0.8cm] {};
+
+        \draw<7-8>[capa] (r2) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,8.75);
+        \draw<7-8>[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,7.5) -- (6.1,10) node (e) [black,midway,xshift=0.8cm] {};
+        \draw<7-8>[capa] (r1) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,11.25);
+        \draw<7-8>[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,10) -- (6.1,12.5) node (e) [black,midway,xshift=0.8cm] {};
+        \node<7-8> at (r3) {0};
+
+        % splice
+        \draw<9>[capa] (r3) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,10);
+        \draw<9>[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,7.5) -- (6.1,12.5) node (e) [black,midway,xshift=0.8cm] {};
+
+        \node<9> at (r1) {0};
+        \node<9> at (r2) {0};
+
+        % splice fail
+        \draw<10->[capa] (r1) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,11.25);
+        \draw<10->[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,10) -- (6.1,12.5) node (e) [black,midway,xshift=0.8cm] {};
+        \draw<10->[capa] (r2) edge[->,postaction={decorate},out=180,in=0,dashed] (6.3,7.75);
+        \draw<10->[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,6.5) -- (6.1,9) node (e) [black,midway,xshift=0.8cm] {};
+
+        \draw<11>[mygreen,very thick] (5.6,10) -- (6.6,9);
+        \draw<11>[mygreen,very thick] (5.6,9) -- (6.6,10);
+        \draw<11> node[mygreen,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, anchor=west] () at (7,9.5) { Splice fails};
+
       \end{tikzpicture}
@@
 \begin{column}{0.4\textwidth}
-  \onslide<3->{Replace pointers with capabilities}
+  \onslide<1->{Linear capabilities}
   \begin{itemize}
-  \item<4-> Pointer
-  \item<5-> Range of authority
-  \item<6-> Permission
-  \end{itemize}
-  \onslide<7->{Dynamically enforced}
-  \begin{itemize}
-  \item<8-> Cap. aware instructions
-  \item<9-> Tagged memory
+  \item<1-> Non-duplicable
+  \item<2-> \texttt{move $r_1$ $r_2$}
+  \item<4-> \texttt{store $r_4$ $r_1$}
+  \item<6-> \texttt{split $r_1$ $r_2$ $r_3$ $n$}
+  \item<8-> \texttt{splice $r_3$ $r_1$ $r_2$}
   \end{itemize}
@@
 \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Call with \stktokens{}}
+  % Emphazise that it is dynamically enforced.
+  % Other cases, see paper.
+  \centering
+  \begin{tikzpicture}[scale=.45, every node={scale=.5}]
+    % recurrent parts
+    \stdstackstart
+    \draw (20,14) node {\phantom{heap memory}};
+
+    \regfile
+
+    % Caller 
+    \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+    \begin{onlyenv}<+>
+      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw[capar] (r1) edge[->,myred,thick,bend left,dashed,postaction={decorate}] (sp1);
+    \end{onlyenv}
+
+    % Split the stack pointer
+    \begin{onlyenv}<+>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw[capar] (r2) edge[->,myred,thick,bend left, dashed,postaction={decorate}] (sp2);
+      \draw node[right=0cm of sp2] { private stack pointer };
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt] (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw[capar] (r1) edge[->,myred,thick,bend left,dashed,postaction={decorate}] (usp);
+      \draw node[right=0cm of usp] { stack pointer };
+    \end{onlyenv}
+
+    % Callee called
+    \begin{onlyenv}<+>
+
+      \draw node[fill=gray,opacity=0.5, right=0cm of sp2] { \phantom{return token} };
+      \draw node[right=0cm of sp2] (rdp) { return token };
+      \draw ($(rdp.west) + (-.1,.3)$) pic {lock};
+      \inactsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \actsf{(0,5)}{(6,8)}{callee stack frame}
+    \end{onlyenv}
+    \begin{onlyenv}<.-.(1)>
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+    \draw[capar] (r2) edge[->,myred,bend left,dashed,postaction={decorate}] (sp2);
+
+    \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt] (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+    \draw[capar] (r1) edge[->,myred,bend left,dashed,postaction={decorate}] (usp);
+    \draw node[right=0cm of usp] { stack pointer };
+    \end{onlyenv}
+
+    % Return from callee
+    \begin{onlyenv}<+>
+      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
+      \draw node[right=0cm of sp2] { private stack pointer };
+    \end{onlyenv}
+
+    \begin{onlyenv}<+>
+      \actsf{(0,2.5)}{(6,5)}{caller stack frame} 
+      \capbrace[sp1]{(6.5,2.5)}{(6.5,16)}
+      \draw node[right=0cm of sp1] { stack pointer };
+      \draw[capar] (r1) edge[->,myred,thick,bend left,dashed,postaction={decorate}] (sp1);
+    \end{onlyenv}
+  \end{tikzpicture}  
 \end{frame}
@@
       \draw node[right=0cm of sp1] { stack pointer };
-      \draw (r1) edge[->,myred,thick,bend left,dashed] (sp1);
+      \draw[capar] (r1) edge[->,myred,thick,bend left,dashed,postaction={decorate}] (sp1);
     \end{onlyenv}
@@
       \capbrace[sp2]{(6.5,2.5)}{(6.5,5)}
-      \draw (r2) edge[->,myred,thick,bend left, dashed] (sp2);
-      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
-      (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,myred,thick,bend left,dashed] (usp);
+      \draw[capar] (r2) edge[->,myred,thick,bend left, dashed,postaction={decorate}] (sp2);
+      \draw [decorate,decoration={brace,aspect=0.2,amplitude=10pt,mirror,raise=4pt},yshift=0pt] (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
+      \draw[capar] (r1) edge[->,myred,thick,bend left,dashed,postaction={decorate}] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
       (6.5,5) -- (6.5,16) node[draw=black] (usp) [black,pos=0.2,xshift=0.8cm] {};
-      \draw (r1) edge[->,myred,bend left,dashed] (usp);
+      \draw[capar] (r1) edge[->,myred,bend left,dashed,postaction={decorate}] (usp);
       \draw node[right=0cm of usp] { stack pointer };
@@
     \end{onlyenv}
-    \draw<.-> (r2) edge[->,myred,bend left,dashed] (sp2);
+    \draw<.->[capar] (r2) edge[->,myred,bend left,dashed,postaction={decorate}] (sp2);
 
@@
 \begin{frame}
-  \frametitle{Paper overview}
+  \frametitle{Full abstraction}
+  \begin{theorem}[Full abstraction]
+  For \tikz[remember picture] \node[inner sep=0pt, outer sep=0pt] (rea) {reasonable};, \tikz[remember picture] \node[inner sep=0pt, outer sep=0pt] (wf) {well-formed};\ $\comp_1$ and $\comp_2$, we have
+  \begin{gather*}
+    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2} \qedhere
+  \end{gather*}
+\end{theorem}
+\vspace{-0.5cm}
+\rule{1\textwidth}{0.4pt}
+\begin{columns}
+  \begin{column}{0.5\textwidth}
+    \begin{itemize}[<3->]
+    \item \stktokens{} reasonable use of return seals
       \begin{itemize}
-\item LCM: A formalization of a simple CHERI-like capability machine with linear capabilities
-\item \stktokens{}, a new calling convention that provably guarantees LSE and WBCF on LCM
-\item A better way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics}
-\item Proof of LSE and WBCF which includes
+      \item Return seals are only used to seal old program pointers.
+      \item Every return seal is only used for one call site.
+      \item Return seals are not leaked.
+      \end{itemize}
+    \item<4-> No measures against us
+  \end{itemize}
+  \end{column}
+  \begin{column}{0.5\textwidth}
+    \begin{itemize}[<6->]
+    \item Component structure.
+    \item Sane linker behaviour.
+    \end{itemize}
+  \end{column}
+\end{columns}
+\vspace{0.5cm}
+\onslide<7>{\centering Should be handled in earlier compiler phase}
+\begin{tikzpicture}[remember picture, overlay]
+  \draw<2-4> node[rectangle, fit=(rea), draw=myred, rounded corners, inner sep=2pt, yshift=-1pt, xshift=0.75pt] (n1) {};
+  \draw<5> node[rectangle, fit=(wf), draw=myred, rounded corners, inner sep=2pt, yshift=-1pt, xshift=0.75pt] (n1) {};
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Conclusion}
 \begin{itemize}
-  \item oLCM: an overlay semantics for LCM with built-in LSE and WBCF
-  \item proving full-abstraction for the embedding of oLCM into LCM by
-  \item defining and using a cross-language, step-indexed, Kripke logical relation with recursive worlds
-  \end{itemize}
+\item \stktokens{} ensures well-bracketed control flow and local-state encapsulation.
+\item Overlay semantics used to define well-bracketed calls and local state encapsulation.
+\item Capability machines may be a good target for secure compilation.
 \end{itemize}
@@
 
+\begin{frame}
+  \frametitle{\stktokens{} vs. CHERI stack}
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      \textbf{\stktokens{}}
+      \begin{itemize}
+      \item No built-in stack
+      \item Fine granularity of security domains
+      \item No context switch on call
+      \item Caller has the full responsibility for securing call
+      \end{itemize}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      \textbf{CHERI stack}
+      \begin{itemize}
+      \item Central, built-in stack
+      \item Call and return instructions interact with stack
+      \item Context switch on call/return
+      \item Local capabilities cannot cross security boundary
+      \end{itemize}
+    \end{column}
+  \end{columns}
+\end{frame}
 \end{document}

### 2019-03-22 presentations/cambridge-18-03-2019/presentation.tex
[Presentation] Update CHERI and StkTokens comparison
[Presentation] Edit a few things, Cambridge 19

@@
       \draw node[right=0cm of sp1] { stack pointer };
-%      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
+      \draw[capa] (r1) edge[->,postaction={decorate},bend left] (sp1);
     \end{onlyenv}
@@
       \draw (16,15) node {heap memory};
+      \node at (r1) {0};
     \end{onlyenv}
@@
     \draw<1-> node[draw=black,rounded corners,above] (source) {High-level language};
-    \draw<1-> node[draw=black,rounded corners,below=of source] (mca) {Machine code w/ many abstractions};
+    \draw<1-> node[draw=black,rounded corners,below=of source] (mca) {Assembly w/ many abstractions};
     \draw<1-> node[draw=black,rounded corners,below=1cm of mca] (mcir) {...};
-    \draw<1-> node[draw=black,rounded corners,below=1cm of mcir] (mcs) {Machine code w/ built-in stack};
-    \draw<1-> node[draw=black,rounded corners,below=of mcs] (target) {Machine code};
+    \draw<1-> node[draw=black,rounded corners,below=1cm of mcir] (mcs) {Assembly w/ built-in stack};
+    \draw<1-> node[draw=black,rounded corners,below=of mcs] (target) {Assembly};
 
@@
     \begin{itemize}[<6->]
-    \item Component structure.
-    \item Sane linker behaviour.
+    \item Component structure
+    \item Sane linker behaviour
     \end{itemize}
@@
   \draw<2-4> node[rectangle, fit=(rea), draw=myred, rounded corners, inner sep=2pt, yshift=-1pt, xshift=0.75pt] (n1) {};
-  \draw<5> node[rectangle, fit=(wf), draw=myred, rounded corners, inner sep=2pt, yshift=-1pt, xshift=0.75pt] (n1) {};
+  \draw<5-6> node[rectangle, fit=(wf), draw=myred, rounded corners, inner sep=2pt, yshift=-1pt, xshift=0.75pt] (n1) {};
 \end{tikzpicture}
@@
 \item \stktokens{} ensures well-bracketed control flow and local-state encapsulation.
-\item Overlay semantics used to define well-bracketed calls and local state encapsulation.
+\item Overlay semantics used to define well-bracketed calls and local-state encapsulation.
 \item Capability machines may be a good target for secure compilation.
@@
       \begin{itemize}
-      \item No built-in stack
+      \item Single stack
       \item Fine granularity of security domains
@@
       \item Caller has the full responsibility for securing call
+      \item Supports higher-order functions
       \end{itemize}
@@
       \begin{itemize}
-      \item Central, built-in stack
+      \item Per component stack and trusted centrally-managed stack
       \item Call and return instructions interact with stack

### 2019-03-26 popl-2019-journal-v/paper.tex
[POPLjv] Add POPL journal version directory
[POPLjv] Add colouring commands and environments
[POPLjv] Add well-formedness definitions and descriptions

--- 2018-11-22 fully-abstract-comp/popl2019/paper.tex
@@
 
-\input{../preamble}
+\input{../fully-abstract-comp/preamble}
 % Paper specific redefinitions of commands
@@
 \setlength{\belowcaptionskip}{-5pt}
+
+\definecolor{Blue}{RGB}{92,26,143}
+\definecolor{OliveGreen}{RGB}{60,128,49}
+\newenvironment{jversion}%{}{}
+    {\color{OliveGreen}}{}
+\newenvironment{jversionsug}
+    {\begin{comment}}{\end{comment}}
+
+\newcommand{\ijversion}[1]%{#1}
+    {{\color{OliveGreen} #1}}
+\newcommand{\ijversionsug}[1]%{}
+    {{\color{Blue} #1}}
 
@@
 % \end{figure}
-\subsection{Well-Formedness and Reasonability}
+\subsection{Well-Formedness}
 \label{sec:well-form-reas}
@@
 The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
-We elide the details of the judgement here and describe the main points:
-For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
-As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
-The return seals $\sigrets$ are the seals supposed to be used to seal return pointers and the closure seals $\sigcloss$ all the other seals in a component.
-If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjoint from the set of trusted addresses $\ta$, then there should be no return seals.
-The code memory may contain sets of seals but only for the seals in $\sigrets$ and $\sigcloss$.
-Other than that, code memory only contains instructions in the form of integers.
-When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), the call must have access to the return seal it specifies.
-The return seal also needs to be unique to that call; that is, no other call can specify the same seal as its return seal.
-The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
-The capabilities in the data memory can only have authority over the data memory itself.
-This allows components to have initial data structures.
-In order to respect Write-XOR-Execute, the capabilities in data memory cannot have execute permission.
-The capabilities also need to respect linearity which means that linear capabilities cannot be aliased by any other capability, and they must be for a range of the predetermined linear addresses $A_\linear$.
-The sealed capabilities must be sealed with a closure seal, and the sealable can be anything that is allowed to reside in data memory.
-The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
-
+The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
+
+\begin{jversion}
+  Informally, a well-formed component respects Write-XOR-Execute.
+  This means that there are no execute capabilities for the data memory and no write capability for the code memory.
+  The data memory of a well-formed component respects linearity and it may come with preset data structures.
+  It may, however, not contain capabilities for the code memory and sets of seals.
+  All the available seals of a component should reside in the code memory.
+  Both adversarial and trusted components can be well-formed.
+  Adversarial components are components where the range of addresses for the code memory is disjoint from the trusted address space Further, adversarial code does not have access to return seals.
+  On the other hand, trusted components must make return seals available to all its calls in a way that is consistent with \stktokens{}.
+  This means that each return seal is only used for one call.
+  Finally, well-formed components can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
+\end{jversion}
+
+% We first highlight the main points of the $\wdjud[\ta]{\comp}$ judgement.
+% For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
+% As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
+% The return seals $\sigrets$ are the seals supposed to be usedto seal return pointers and the closure seals $\sigcloss$ all the other seals in a component.
+% If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjoint from the set of trusted addresses $\ta$, then there should be no return seals.
+% The code memory may contain sets of seals but only for the seals in $\sigrets$ and $\sigcloss$.
+% Other than that, code memory only contains instructions in the form of integers.
+% When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), the call must have access to the return seal it specifies.
+% The return seal also needs to be unique to that call; that is, no other call can specify the same seal as its return seal.
+% The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
+% The capabilities in the data memory can only have authority over the data memory itself.
+% This allows components to have initial data structures.
+% In order to respect Write-XOR-Execute, the capabilities in data memory cannot have execute permission.
+% The capabilities also need to respect linearity which means that linear capabilities cannot be aliased by any other capability, and they must be for a range of the predetermined linear addresses $A_\linear$.
+% The sealed capabilities must be sealed with a closure seal, and the sealable can be anything that is allowed to reside in data memory.
+% The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
+
+\begin{jversion}
+The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
+The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
+
+%TODO distinguish closure seals from return seals.
+The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
+The \textsc{Main} rule requires the component to have a main pair picked from the exports.
+Further, the component should be a base component that satisfies judgement.
+The \textsc{Base} judgement makes sure that the component has a bit of structure.
+Roughly speaking, the \textsc{Base} rule requires the following:
+\begin{itemize}
+\item \textit{Code and data memory are disjoint.}
+\item \textit{Code memory is enclosed in a memory segment $\ms_\var{pad}$ that contains zero's.} This prevents the memory of two code memories from being spliced together.
+\item \textit{Data memory can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.} The component judgement ensure that the linear address space of a component is at most referenced by one capability.
+This means that any capability for the linear address space is unique.
+\item \textit{There is a partition $A_\var{own}$ of $A_\var{linear}$ that assigns a set of linear-addressed to each address of data memory.} The partitions are used by the component word judgement ($\vdash_\mathrm{comp-word}$) which ensures linear capabilities only address the addresses in the linear address space that is owned by the address where the capability resides in memory.
+\item \textit{All import addresses are part of the data memory.} 
+\item \textit{Import and export symbols are disjoint.}
+\item \textit{One of the following two are true}
+  \begin{itemize}
+  \item \textit{The code memory is disjoint from the trusted address space $\ta$ and in that case there are no return seals.} In this case, the component contains untrusted code.
+We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals.
+  \item \textit{The code memory is part of the trusted address space $\ta$.} In this case, the component contains trusted code.
+We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
+  \end{itemize}
+\item \emph{The data memory is disjoint from the trusted address space $\ta$.} We are not looking to execute data, so the trusted addresses never include the data memory addresses.
+\item \textit{The code memory satisfies the component-code judgement (\figurename~\ref{fig:well-formed-code}).}
+\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).
+In particular, this must be with respect to the linear addresses assigned to this address.
+}
+\item \textit{All the exports satisfies the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
+\end{itemize}
+
+\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ the words that are well-formed in a component's data memory.
+The judgement is defined by three rules: \textsc{W-Data}, \textsc{W-Capability}, and \textsc{W-Sealed-Capability}.
+The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
+The \textsc{W-Capability} rule specifies that all well-formed capabilities in data memory must at most have read and write permission.
+Further, if the capability is linear, then the range of authority must be within the linear address space owned by this address.
+Finally, if it is non-linear, then it must be in the non-linear address space.
+The \textsc{W-Sealed-Capability} specifies that well-formed sealed capabilities in data memory are sealed with a closure seal and whatever is sealed is itself a well-formed word.
+Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
+However, a component is free to place sealed data capabilities in memory during execution.
+
+\figurename~\ref{fig:well-formed-export} defines the well-formed exports $\vdash_{\mathrm{comp-export}}$.
+\textsc{E-Word} says that any well-formed word for data memory is well-formed as an export.
+In particular, this means that sealed capabilities for the data memory are acceptable exports which allows components to export the data part of a sealed capability pair.
+The \textsc{E-Sealed-Code} rule allows the code part of a sealed capability pair to be export.
+In particular, it allows a capability with read-execute permission and its range of authority within the component's code address space to be exported when it has been sealed with a closure seal.
+Together, these two rules allow components to have closures that they export.
+
+\figurename~\ref{fig:well-formed-code} defines the well-formed code memories $\vdash_{\mathrm{comp-code}}$.
+The well-formed code memories are defined by two judgements.
+One judgement considers the entire memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ while the other considers the contents of each code memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
+Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each of its words in isolation.
+The well-formedness of code memory depends on whether seals for calls are located in the place we expect them to be located.
+The $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$ judgement has two rules.
+The \textsc{C-Seal} judgement requires address $a$ in $\mscode$ to contain a set of seals that is a subset of the closure seals and the return seals.
+\textsc{C-Instr} requires address $a$ to contain an integer, which may be interpreted as instruction.
+If address $a$ is the first address of a series of instructions that may be interpreted as a call, and all the addresses are within the trusted address space, then there must be a set of seals available at the address the call in question expects it and that set of seals must contain the return seal that corresponds to the return seal of this call.
+The return seal in question must be in the $\sigrets[\mathrm{owned}]$ to ensure that no other call uses it.
+In order to ensure that return seals are at most used by one call, the $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ judgement takes all the available return seals and partitions them over code addresses.
+This means that no other call can use the same seal.
+Further, return seals cannot be used to seal non-return capabilities, so the \textsc{C-Mem} judgement requires the set of return seals to be disjoint from the set of closure seals.
+To limit the amount of corner cases we have to consider, we require the code memory to have no \textit{hidden calls}.
+\lau{Do we actually use the hidden calls in the proofs?
+I vaguely recall this was added to make sure that splicing would not create a call.
+However, we also have the padding which, at least in part, takes care of this.}
+\begin{definition}[No hidden calls]
+  \label{def:no-hidden-calls}
+  We say that a memory segment $\mscode$ has no hidden calls iff
+  \[
+    \begin{array}{l}
+      \forall \aaddr \in \dom(\mscode) \ldotp \\
+      \quad \forall i \in [0,\calllen - 1] \\
+      \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
+      \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1] \wedge\\
+      \qquad \qquad \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
+      \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
+      \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
+    \end{array}
+  \]
+\end{definition}
+This definition says that if code memory contains an instruction that may be part of a call, then either it is a part of a call or there is some other instruction that witnesses the fact that it is not part of a call.
+Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set of seals available.
+
+\begin{figure}
+\begin{mathpar}
+  \inferrule*[Right=Base]{
+    \dom(\mscode) = [\baddr,\eaddr]\\
+    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
+    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
+    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}\\
+    \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
+    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
+    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} \\
+    \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}} \\
+    \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
+    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} \\
+    (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset) \\
+    \dom(\msdata) \mathrel{\#} \ta \\
+    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
+    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
+    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
+  }{
+    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+  }
+  \and
+  \inferrule*[Right=Main]{
+    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
+    \ta \vdash \var{comp}_0 \\
+    (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
+  }{
+    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
+  }
+\end{mathpar}
+\caption{Well-formedness judgement.}
+\label{fig:well-formed}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{subfigure}[h]{1.0\linewidth}
+    \begin{mathpar}
+  \inferrule*[Right=C-Seals]{
+    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \\ [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
+  }{
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
+  }
+  \and
+  \inferrule*[Right=C-Instr]{
+    \mscode(a) \in \ints\\
+{    \begin{multlined}
+      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
+      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
+    \end{multlined}
+}  }{
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
+  }
+  \and
+  \inferrule*[Right=C-Mem]{
+    \text{$\mscode$ has no hidden calls}\\
+    \sigrets \mathrel{\#} \sigcloss \\
+    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
+    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
+    \forall a \in \dom(\mscode) \ldotp 
+    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
+    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
+  }{
+    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
+  }
+    \end{mathpar}
+    \caption{Code well-formedness.}
+    \label{fig:well-formed-code}
+  \end{subfigure}
+
+  \begin{subfigure}[h]{1.0\linewidth}
+    \begin{mathpar}
+      \inferrule*[Right=W-Data]{
+        z \in \ints
+  }{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} z
+  }
+  \and
+  \inferrule*[Right=W-Capability]{
+    \permbnf \sqsubseteq \rw\\
+    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}} \\\\
+    \lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
+  }{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+  }
+  \and
+  \inferrule*[Right=W-Sealed-Capability]{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \vsc \\
+    \sigma \in \sigcloss
+  }{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \sealed{\sigma,\vsc}
+  }
+    \end{mathpar}
+    \caption{Word well-formedness.}
+    \label{fig:well-formed-word}
+  \end{subfigure}
+
+  \begin{subfigure}[h]{1.0\linewidth}
+    \begin{mathpar}
+  \inferrule*[Right=E-Sealed-Code]{
+    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
+    \sigma \in \sigcloss
+  }{
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
+  }
+  \and
+  \inferrule*[Right=E-Word]{
+    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
+  }{
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
+  }
+    \end{mathpar}
+    \caption{Export well-formedness}
+    \label{fig:well-formed-export}
+  \end{subfigure}
+  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
+  \label{fig:well-formed-aux}
+\end{figure}
+\end{jversion}
+\subsection{Reasonability}
 % Reasonability
@@
              \left\{\begin{multlined}
-               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg}) \times \\
+               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
                (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})

### 2019-03-26 popl-2019-journal-v/paper.tex
[POPLjv] Edit minor things in the text

@@
 \todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
+Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be built.
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
 Generally, properties that hold in the source language but not in the target language need to be somehow enforced by the compiler.
-Two properties that hold in many high-level source languages, but not in the assembly languages they are compiled to, are well-bracketed control flow and encapsulation of local state.
+Two properties that hold in many high-level source languages are well-bracketed control flow and encapsulation of local state, but they are not enforced after compilation to assembly.
 
@@
 The hardware prevents these capabilities from being duplicated.
-We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
+We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to give them back on return.
 The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
@@
 
-\begin{wrapfigure}{r}{0.19\linewidth}
+\begin{wrapfigure}{r}{0.30\linewidth}
   \centering
@@
 % seals
+A sealed capability $\sealed{\sigma,\vsc}$ is a pair of a seal $\sigma$ and a capability $\vsc$.
 A sealed capability makes an underlying capability opaque which means that the underlying capability cannot be changed or used for the operations it normally gives permission to.
 In other words, the authority the underlying capability represents is encapsulated under the seal.
-Syntactically, the sealed capability is represented as $\sealed{\sigma,\vsc}$ where $\vsc$ is a sealable and $\sigma$ is the seal used to seal the capability.
-In order to seal a sealable with a seal $\sigma$, it is necessary to have the right to do so.
-The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
+In order to seal a sealable with a seal $\sigma$, it is necessary to have the permission to do so.
+The permission to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
 A set of seals is a capability that represents the authority to seal sealables with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In spirit of memory capabilities, a set of seals has an active seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
+In spirit of memory capabilities, a set of seals has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
 % On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
@@
 %%% Domains
-Words on \trgcm{} are capabilities and data (represented by integers ($\ints$)).
-We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
+Words on \trgcm{} are capabilities and data (represented by integers $\ints$).
+We assume a finite set of register names $\RegName$ containing at least the registers $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$.
+We define register files as functions from register names to words.
 Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ partial functions).
@@
 It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
-When a linear capability is moved, the source register must be cleared, so the capability is not duplicated.
+When a linear capability is moved, the source register must be cleared to prevent duplication of the capability.
 This is done uniformly in the semantics using the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.

### 2019-03-29 popl-2019-journal-v/paper.tex
[POPLjv] Add reasonable word definition and subsubsection

@@
 % \end{figure}
-\subsection{Well-Formedness}
+\subsection{Well-Formed Components}
 \label{sec:well-form-reas}
@@
 \end{jversion}
-\subsection{Reasonability}
+\subsection{Reasonable Components}
 % Reasonability
@@
 In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
-More precisely, we expect four things of a reasonable component: (1) It checks the stack base address before performing a call.
+More precisely, we expect four things of a reasonable component:
+\begin{enumerate}[label=(\arabic*)]
+\item It checks the stack base address before performing a call.
 As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
-(2) It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
-(3) It does not leak return and closure seals or means to retrieve them.
+\item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
+\item It does not leak return and closure seals or means to retrieve them.
 This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
 There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
-(4) It does not store return and closure seals or means to get them.
+\item It does not store return and closure seals or means to get them.
 By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
-We elide the details here and refer to~\citet{technical_report}
+\end{enumerate}
+\begin{jversion}
+  We capture these properties in 4 definitions.
+  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly as well as indirectly.
+  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then  the configuration behaves reasonable.
+  Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
+  Finally, definition~\ref{def:reasonable-component} lifts the notion of reasonability to components.
+
+  \subsubsection{Reasonable words}
+  To provide any guarantees, \stktokens{} rely on the program to no leak return seals in any way.
+  But what does it mean to ``leak'' a return seal.
+  This means that sets of seals that contain return seals as well as any means to obtain such sets cannot be leaked.
+  The following definition makes this more precise.
+\begin{definition}[Reasonable word]
+  \label{def:reasonable-word}
+  Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
+We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
+free stack $\ms_\stk$ if $n=0$
+or the following implications hold.
+  \begin{itemize}
+  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then
+    $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \mathrel{\#} \dom(\ta)$
+  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
+    \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$
+    and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all
+    $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
+  \item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$
+    and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for
+    all $\aaddr \in [\baddr,\eaddr]$
+  \end{itemize}
+\end{definition}
+According to the definition, sets of seals that contain return or closure seals are not reasonable.
+\stktokens{} need the return seals to be unique in order to work, but it does not need the closure seals.
+However, it defeats the purpose of the closure seals if they are leaked to malicious code as it allows the malicious code to fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectually allows malicious code to unseal the data capability.
+Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed in the context of.
+
+% TODO Mention sealed capabilities?
+
+A reasonable word is defined relative to a memory and a stack because for capabilities it depends on what they give access to whether they are reasonable.
+For instance, if a read capability gave access to a piece of memory that contains a set of seals with a return seal, then it would not be reasonable.
+This is why stack pointers and memory capabilities with read permission must only give access to memory with reasonable words.
+
+The definition is cyclic, but the step-index ensures that it is well-founded.
 
@@
 % Reasonable component
+
+\end{jversion}
 

### 2019-04-01 popl-2019-journal-v/paper.tex
[POPLjv] Insert reasonability section with definitions and text

@@
 
-In our result, we assume that adversarial components are well-formed, but not necessarily reasonable.
-The well-formedness assumption ensures that the trusted component can rely on basic security guarantees provided by the capability machine.
+\subsubsection{Reasonable pc and configuration}
+In order to define desired behaviour, we need to say specify what may happen on a machine during execution.
+The execution steps between executable configurations, so we need to define when a configuration is reasonable.
+While the step relation is defined over configurations, it is the pc that decides what instruction is executed.
+It therefore only seem natural to also define when a pc is reasonable.
+Definition~\ref{def:reasonable-pc} roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file is reasonable and the result is a reasonable configuration.
+  \begin{definition}[Reasonable pc]
+    \label{def:reasonable-pc}
+    We say that an executable capability $c =
+    ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps
+    if for any $\Phi$ such that
+    \begin{itemize}
+    \item $\Phi.\reg(\pcreg) = c$
+    \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and
+      free stack $\Phi.\ms_\stk$ for all $r \neq \pcreg$
+    \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
+    \end{itemize}
+    We have that $\Phi$ is reasonable up to $n$ steps.
+  \end{definition}
+  With Definition~\ref{def:reasonable-word} and \ref{def:reaonable-pc} in place, we are all set to define when a configuration is reasonable.
+  \begin{definition}[Reasonable configuration]
+    \label{def:reasonable-conf}
+    We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
+    iff for $n' \leq n$:
+    \begin{enumerate}
+    \item \label{item:guarantee-stk-base} \emph{Guarantee stack base address before call...} \\
+      If $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$
+      for some $\src{r_1}$ and $\src{r_2}$,
+      then all of the following hold:
+      \begin{itemize}
+      \item $\src{\Phi}(\src{r_\stk}) =
+        \src{\stkptr{\_,\stkb,\_,\_}}$
+      \item $r_1 \neq \rtmp{1}$
+      \item $n' = 0$ or $\Phi(\pcreg) + \calllen$ behaves reasonably up to $n'-1$ steps (Definition~\ref{def:reasonable-pc})
+      \end{itemize}
+    \item \label{item:return-seal-usage} \emph{Use return seals only for calls, use closure seals appropriately...} \\
+      If $\src{\Phi}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$,
+      then one of the following holds:
+      \begin{itemize}
+      \item $\src{\Phi}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$
+      \item $\sigma \in \gsigcloss$ and one of the following holds:
+        \begin{itemize}
+        \item $\exec{\Phi(r_1)}$ and $n' = 0$ or $\Phi(r_1)$ behaves reasonably up to $n' - 1$ steps (Definition~\ref{def:reasonable-pc}).
+        \item $\nonExec{\Phi(r_1)}$ and $n' = 0$ or $\Phi(r_1)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$ (Definition~\ref{def:reasonable-word}).
+        \end{itemize}
+      \end{itemize}
+    \item \label{item:dont-store-private}\emph{Don't store private stuff...}\\
+      If $\src{\Phi}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$,
+      then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ steps.
+    \item \label{item:dont-leak-private}\emph{Don't leak private stuff...}\\
+      If $\Phi \step[\ta,\stkb] \Phi'$, then one of the following holds:
+      \begin{itemize}
+      \item All of the following hold:
+        \begin{itemize}
+        \item $\Phi'.\reg(\pcreg) =
+          ((\perm,\lin),\baddr,\eaddr,\aaddr')$ and $\Phi.\reg(\pcreg) =
+          ((\perm,\lin),\baddr,\eaddr,\aaddr)$
+        \item $\Phi$ does not point to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+        \item $\Phi$ does not point to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$, $\offpc$, $\offsigma$
+        \item $n' = 0$ or $\Phi'$ is reasonable up to $n'-1$ steps
+        \end{itemize}
+        % \dominique{the following seems unnecessary: why would we allow trusted code to use a jump instruction to jump out of a trusted component?
+it can use call and xjmp for calls/returns/tailcalls}
+        % \item $\Phi'.\reg(r)$ is reasonable in memory $\Phi'.\mem$ and free stack $\Phi'.\ms_\stk$ up to $n'-1$ steps for all $r$ (including $\pcreg$)
+      \item All of the following hold:
+        \begin{itemize}
+        \item $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+        \item $n' = 0$ or $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
+        \end{itemize}
+      \item All of the following hold:
+        \begin{itemize}
+        \item $\Phi$ points to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
+        \item $n' = 0$ or $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
+        \end{itemize}
+      \end{itemize}
+    \end{enumerate}
+  \end{definition}
+The four items in Definition~\ref{def:reasonable-conf} correspond to the four informal items from the introduction of this section.
+
+Item~\ref{item:dont-store-private} and Item~\ref{item:don't-leak-private} makes sure that return seals are not leaked.
+Item~\ref{item:dont-store-private} says that only reasonable words can be stored to memory.
+This means that sets of return seals or execute capabilities cannot be stored to memory by a reasonable component.
+Intuitively, a well-formed and reasonable component has its seals available in the code memory, so it can always retrieve them from the code memory.
+In other words, it is not necessary to store them in the data memory.
+Further, by making sure that seals are not stored to memory, we can allow capabilities for data memory to be passed away if there is a need for that (for instance to have a shared buffer).
+Item~\ref{item:dont-leak-private} makes sure that seals are not leaked when transferring control to another component (i.e. on security boundary crossings).
+With the component setup, there are two ways to transfer control: xjmp and call.
+In both cases, we require that all of the argument registers contain reasonable words.
+An execution configuration may need to do other operations than calling other code, and seals should no be leaked at any point during execution.
+For this reason, Item~\ref{item:don-leak-private} also says that if the next step is not a call and a jump, then the next execution configuration should also be reasonable.
+
+Item~\ref{item:guarantee-stk-base} makes sure that the stack has the correct base before a call.
+In order to not have to reason about unreasonably generated code, we also add the requirement $r_1 \neq \rtmp{1}$.
+If we allowed $r_1=\rtmp{1}$, then the call would be sure to fail as the first instruction of a call moves 42 to $\rtmp{1}$.
+Finally, this promises that the code after the call will behave reasonably.
+
+A well-formed component makes sure that a return seal is uniquely available to every call.
+This is, however, not sufficient as it does not ensure that other parts of a program don't use the return seals.
+We do not want to specify what non-call code should look like, so we just require it to not use the call seals.
+This is what Item~\ref{item:return-seal-usage} ensures.
+It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction is used to seal part of a closure which means that it should be using one of the closure seals.
+In order to construct a closure, one must seal a code capability and a data capability.
+If this is an executable capability, then it should be reasonable as a pc.
+On the other hand, if it is not an executable capability, then this must be the data part of the pair, so it should just be a reasonable word.
+
+Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle. If
+
+  \subsubsection{Reasonable component}
+  A reasonable component has the informal behavioural properties from the introduction.
+  Reasonability is captured by the previous definitions. These definitions are lifted to components by the following definition.
+  \begin{definition}[Reasonable component]
+    \label{def:reasonable-component}
+    We say that a component
+    $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$
+    is reasonable if the following hold: For all $(s \mapsto \sealed{\sigma,\vsc}) \in
+    \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$, we have that $\vsc$
+    behaves reasonably up to any number of steps $n$.
+
+    We say that a component
+    $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is reasonable if $\var{comp}_0$ is reasonable.
+    % \lau{Is this reasonable as a word? In that case, it should  mention the
+    %   memory and free stack.}
+    % \dominique{No: this is reasonable as a  pc.}
+  \end{definition}
+\end{jversion}
+
+In our result, we assume that adversarial components are well-formed, but not necessarily reasonable. The well-formedness assumption ensures
+ that the trusted component can rely on basic security guarantees provided by the capability machine.
 For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
@@
 We will assume trusted components, for which WBCF and LSE are guaranteed, to be both well-formed and reasonable.
-%The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.
-
-% \begin{definition}[Reasonable configuration]
-%   \label{def:reasonable-conf}
-%   We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
-%   iff for $n' \leq n$:
-%   \begin{itemize}
-%   \item \emph{Guarantee stack base address before call...}
-%   \item \emph{Use return seals only for calls, use closure seals appropriately...}
-%   \item \emph{Don't store private stuff...} If $\src{\Phi}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
-
-%     Then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ steps.
-%   \item \emph{Don't leak private stuff...}
-%   \end{itemize}
-% \end{definition}
-% Reasonable configuration
-% Reasonable component
-
-\end{jversion}
 

### 2019-04-01 popl-2019-journal-v/paper.tex
[POPLjv] Add text about decode and encoding functions

@@
 \renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
+\renewcommand{\encInstr}[1]{\plainfun{encode}{#1}}
 \renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
@@
 
+\begin{jversion}
+\subsection{Decoding and encoding functions}
+The operational semantics of the capability machine uses $\decInstr{}$ to decode instructions.
+We also assume an $\encInstr{}$ which is used to put programs onto the capability machine.
+Rather than specifying a decode function and an encode function, we assume that they are given with certain properties.
+The $\decInstr{} : \Word \fun \Instr$ should be injective and surjective for all non-$\fail$ instructions.
+Further, it should decode all capabilities as the $\fail$ instruction, i.e.\
+\[
+\forall c \in \Cap \ldotp \decInstr{c} = \fail
+\]
+The $\encInstr{} : \Instr \fun \ints$ function should be injective, and it should be defined so $\decInstr{}$ is its left inverse, i.e.
+\[
+\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
+\]
+
+These assumptions are sufficient to construct program examples and run them on the machine.
+The $\encInstr{}$ function allows us to specify the program ``abstractly'' in terms of instructions rather than machine words.
+As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
+%This is not to say that encoding is an easy problem. There are many things to get.
+
+The machine also assumes decode and encode functions for permissions $\decPerm{} : \Perm \fun \ints$ and $\encPerm{} : \ints \fun \Perm$.
+We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$ and surjective. For $\encPerm{}$, we assume it is surjective and that it does not encode anything to the error value -1, i.e.
+\[
+  \forall \perm \in \Perm \ldotp \encPerm{\perm} \neq -1
+\]
+For linearity we assume similar functions.
+
+Finally, the machine uses an encode function for word types $\encType{}$ in the interpretation of $\tisptr{}{}$.
+This function encodes each kind type of word as an integer.
+This is very much like the previous functions.
+It encodes each kind of word differently and all words of the same kind to the same integer.
+In Section~\ref{subsec:overlay-semantics}, we present an overlay semantics which changes the set of words compared to this machine.
+The overlay semantics also has a type encoding function which must encode the new words, so they correspond to their representation on this machine.
+For instance, the encoding of the type of a stack-pointer in the overlay semantics should be the same as the encoding of a capability on this machine.
+All the new words correspond to a word on this machine, and their encodings should match.
+\end{jversion}
+
+\subsection{Components, linking, programs, and contexts }
 The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.

### 2019-04-02 popl-2019-journal-v/paper.tex
[POPLjv] Add example to motivate the necessity of sealing

@@
 %%% Seals
-% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
 
@@
 In other words, the authority the underlying capability represents is encapsulated under the seal.
-In order to seal a sealable with a seal $\sigma$, it is necessary to have the permission to do so.
+In order to seal a sealable with a seal $\sigma$, it is necessary to have the authority to do so.
 The permission to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
@@
 This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.
+
+\begin{jversion}
+  \subsection{The purpose of sealing}
+  % Sealing example:
+  To motivate the necessity of an encapsulation mechanism like sealing, consider the following example.
+  We want to transfer control to code that we do not trust, and we want to give them the means to return to us.
+  That is, we need to give them a return capability.
+  If we did not have an encapsulation mechanism, our only option would be to give them an executable capability for the address we want them to return to.
+  The code we give the return capability could use it as intended, but it could also manipulate the return capability to point to a different point and return to an unintended point of our code.
+  Further in order to be able to retrieve our capabilities from before transferring control, we would have to store them somewhere accessible from the return capability.
+  However, the untrusted code would have access to all this through the return capability because it does not give less authority to the untrusted code than it does to us.
+  For this reason, any reasonable capability machine must have an encapsulation mechanism to enable programs to make boundaries between security domains.
+
+  In the example, if we use sealing, we can seal the return capability which does two things.
+  It prevents the untrusted code from changing the target of the return capability forcing them to return to the point we intended.
+  It also prevents the untrusted code from reading our capabilities.
+  All in all, this means that we can transfer control to another domain without giving up our capabilities or handing them over.
+\end{jversion}
 
@@
 \subsection{Overlay Semantics}
+\label{subsec:overlay-semantics}
 % Describe the things added to the syntax

### 2019-04-04 popl-2019-journal-v/paper.tex
[POPLjv] Describe unsealing

@@
   All in all, this means that we can transfer control to another domain without giving up our capabilities or handing them over.
+
+  \trgcm{} does not have a direct unsealing instruction, but it is still possible to do some amount of unsealing.
+  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains the $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
+  Now assume part of the code would like to have access to $w$ in $\rdata$.
+  Take a capability for this code, possibly by adjusting the pc, and seal it with $\sigma$.
+  Now you have the data part and the code part of a sealed capability pair, so you can $\txjmp{}{}$ to it which unseals the data capability and puts it in $\rdata$ for your code to use.
+  Sealed executable capabilities cannot be unsealed in the same way because $\txjmp{}{}$ fails if the data capability is executable.
+  However, given a sealed executable capability $\sealed{\sigma,c}$ and a set of seals that contains $\sigma$, we can still construct the data part of the sealed capability pair.
+  This means that we can execute the code the capability points to in the presence of any data we can construct.
+
+  Sealing is meant for encapsulation, but it gives us very weak encapsulation properties if seals are not private as the unsealing example illustrates.
+  For this reason, it is important that the system is setup, so each component has access to seals that are unique to that component.
+  We return to this in \sectionname~\ref{sec:well-form-reas}.
 \end{jversion}

### 2019-04-04 popl-2019-journal-v/appendix.tex
[POPLjv] Add missing instructions to appendix

\section{\trgcm{} instruction interpretation}
\label{app:instr-interpretation}
In this section, we present the interpretation of the \trgcm{} instructions left out of \figurename~\ref{fig:target-op-sem} as well as the interpretation of the \srccm{} specific interpretation.
The two interpretations have everything in black in common, and everything in blue is specific to \srccm{}.

 \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.2cm}|}
    \hline
    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
    \hline
    \tgetb{r_1}{r_2}                                        & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\baddr,\_,\_)$, $\Phi(r_2) = \seal{\baddr,\_,\_}$, \src{or $\Phi(r_2) = ((\_,\_),\baddr,\_,\_)$}, then $w = \baddr$ and otherwise $w = -1$\\
    \hline
    \tgete{r_1}{r_2}                                        & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\eaddr,\_)$, or $\Phi(r_2) = \seal{\_,\eaddr,\_}$, \src{or $\Phi(r_2) = ((\_,\_),\_,\eaddr,\_)$}, then $w = \eaddr$ and otherwise $w = -1$\\
    \hline
    \tisptr{r_1}{r_2} & \updPcAddr{\Phi\updReg{r_1}{w}} & $w = \encType{\Phi(r_2)}$\footnote{The $\encType$ function differs for the two machines. Depending on the machine, the appropriate function is used.}\\
    \hline
    \tgetlin{r_1}{r_2} &\updPcAddr{\Phi\updReg{r_1}{w}} & If $\isLinear{\Phi(r_2)}$, then $w = \encLin{\linear}$, otherwise $w= \encLin{\normal}$\\
    \hline
    \tgetp{r_1}{r_2} & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\perm,\_),\_,\_,\_)$ \src{or $\Phi(r_2) = ((\perm,\_),\_,\_,\_)$}, then $w = \encPerm{\perm}$ and otherwise $w = -1$\\
    \hline
    \tjnz{r}{\rn} & \Phi\updReg{r}{w}\update{\pcreg}{\Phi(r)} & $\nonZero{\Phi(\rn)}$ and $w = \linCons{\Phi(r)}$\\
    \hline
    \tjnz{r}{\rn} & \updPcAddr{\Phi}& If not $\nonZero{\Phi(\rn)}$ \\
    \hline
    \tplus{r}{\rn_1}{\rn_2} & \updPcAddr{\Phi\updReg{r}{n_1+n_2}} & If for $i \in \{1,2\}$ $\Phi(\rn_i) = n_i \in \ints$ \\
    \hline
    \tminus{r}{\rn_1}{\rn_2} & \updPcAddr{\Phi\updReg{r}{n_1-n_2}} & If for $i \in \{1,2\}$ $\Phi(\rn_i) = n_i \in \ints$ \\
    \hline
    \tlt{r}{\rn_1}{\rn_2} & \updPcAddr{\Phi\updReg{r}{1}} & If for $i \in \{1,2\}$ $\Phi(\rn_i) = n_i \in \ints$ and $n_1 < n_2$ \\
    \hline
    \tlt{r}{\rn_1}{\rn_2} & \updPcAddr{\Phi\updReg{r}{0}} & If for $i \in \{1,2\}$ $\Phi(\rn_i) = n_i \in \ints$ and $n_1 \not< n_2$ \\
    \hline
    \tsetatob{r} & \updPcAddr{\Phi\updReg{r}{c}} & $r \neq \pcreg$, $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\_)$, and $c = ((\perm,\lin),\baddr,\eaddr,\baddr)$ \\
    \hline
    \tsetatob{r} & \updPcAddr{\Phi\updReg{r}{c}} & $r \neq \pcreg$, $\Phi(r) = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, and $c = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ \\
    \hline
    \src{\texttt{seta2b}\; r} & \src{\updPcAddr{\Phi\updReg{r}{c}}} &\srcalt{$r \neq \pcreg$, $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\_}$, and $c = \stkptr{\perm,\baddr,\eaddr,\baddr}$} \\
    \hline
    \trestrict{r}{\rn} & \updPcAddr{\Phi\updReg{r}{c}} & If $\Phi(r) = ((\perm,\lin),\baddr, \eaddr,\aaddr)$ and $\Phi(\rn) = n$ and $\decPerm{n} \le \perm$ and $c = ((\decPerm{n},\lin),\baddr, \eaddr, \aaddr)$\\
    \hline
    \src{\texttt{restrict} \; r \; \rn} & \src{\updPcAddr{\Phi\updReg{r}{c}}} & \srcalt{If $\Phi(r) = \stkptr{\perm,\baddr, \eaddr,\aaddr}$ and $\Phi(\rn) = n$ and $\decPerm{n} \le \perm$ and $c = \stkptr{\decPerm{n},lin),\baddr, \eaddr, \aaddr}$}\\
    \hline
  \end{tabular}

Where $\nonZero{}$ is defined as follows
\[
  \nonZero{w} \defeq
  \begin{cases}
    \bot & w \in \ints \tand w = 0 \\
    \top & \totherwise
  \end{cases}
\]

%%% Local Variables:
%%% TeX-master: "paper"
%%% End:

### 2019-04-04 popl-2019-journal-v/paper.tex
[POPLjv] Add split oLCM interpretation

@@
 \begin{jversion}
+% The remaining instructions
+The interpretation of the remainder of the instructions are displayed in Appendix~\ref{app:instr-interpretation}.
+The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all project information about capabilities.
+The $\tgetb{}{}$ instruction projects the base address of the range of authority, and $\tgete{}{}$ projects the end address of the range of authority.
+The linearity of a permission is projected with $\tgetlin{}{}$, and finally the permission is projected with $\tgetp{}{}$.
+The instructions $\tgetb{}{}$ and $\tgete{}{}$ also works on sets of seals.
+The instruction $\tisptr{}{}$ gives the type of a word which allows programs to be defensive in the sense that they can check whether a word is a capability before they use the word as one.
+\trgcm{} also has arithmetic instructions $\tplus{}{}{}$, $\tminus{}{}{}$, and $\tlt{}{}{}$.
+The latter instruction compares two numbers and writes 1 or 0 to a target register depending on whether one number is less than the other.
+The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set of seals) to the base address of the range of authority (or range of seals).
+This instruction makes it possible to work relatively to the base address of a capability rather than with fixed addresses.
+While this instruction is not strictly necessary as it can be emulated with other instruction, it still seems useful in the case where you receive capabilities from an untrusted source.
+Finally, we have the $\trestrict{}{}$ instruction which  restricts the permission of a capability but only according to the $\le$ relation.
+\end{jversion}
+
+\begin{jversion}
   \subsection{The purpose of sealing}
@@
 \subsection{Components, linking, programs, and contexts }
+\label{subsed:components-linking}
 The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
@@
     \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
+    \hline
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & $\sourcecolor\left.\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{0} \\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray\right.$ & \srcalt{$\Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = \stkptr{\perm,\baddr,n,\aaddr}$ and $c_2 = \stkptr{\perm,n+1,\eaddr,\aaddr}$}\\
     \hline
@@
 
+\clearpage
+\appendix
+\input{appendix}
 \end{document}

### 2019-04-05 popl-2019-journal-v/appendix.tex
[POPLjv] Add oLCM missing interpretations to appendix

@@
 \label{app:instr-interpretation}
-In this section, we present the interpretation of the \trgcm{} instructions left out of \figurename~\ref{fig:target-op-sem} as well as the interpretation of the \srccm{} specific interpretation.
-The two interpretations have everything in black in common, and everything in blue is specific to \srccm{}.
+In this section, we present the interpretation of the \trgcm{} instructions left out of \figurename~\ref{fig:target-op-sem}.
 
@@
     \hline
-    \tgetb{r_1}{r_2}                                        & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\baddr,\_,\_)$, $\Phi(r_2) = \seal{\baddr,\_,\_}$, \src{or $\Phi(r_2) = ((\_,\_),\baddr,\_,\_)$}, then $w = \baddr$ and otherwise $w = -1$\\
+    \tgetb{r_1}{r_2}                                        & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\baddr,\_,\_)$ or$\Phi(r_2) = \seal{\baddr,\_,\_}$, then $w = \baddr$ and otherwise $w = -1$\\
     \hline
-    \tgete{r_1}{r_2}                                        & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\eaddr,\_)$, or $\Phi(r_2) = \seal{\_,\eaddr,\_}$, \src{or $\Phi(r_2) = ((\_,\_),\_,\eaddr,\_)$}, then $w = \eaddr$ and otherwise $w = -1$\\
+    \tgete{r_1}{r_2}                                        & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\eaddr,\_)$ or $\Phi(r_2) = \seal{\_,\eaddr,\_}$, then $w = \eaddr$ and otherwise $w = -1$\\
     \hline
@@
     \hline
-    \tgetp{r_1}{r_2} & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\perm,\_),\_,\_,\_)$ \src{or $\Phi(r_2) = ((\perm,\_),\_,\_,\_)$}, then $w = \encPerm{\perm}$ and otherwise $w = -1$\\
+    \tgetp{r_1}{r_2} & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\perm,\_),\_,\_,\_)$, then $w = \encPerm{\perm}$ and otherwise $w = -1$\\
     \hline
@@
     \hline
-    \src{\texttt{seta2b}\; r} & \src{\updPcAddr{\Phi\updReg{r}{c}}} &\srcalt{$r \neq \pcreg$, $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\_}$, and $c = \stkptr{\perm,\baddr,\eaddr,\baddr}$} \\
-    \hline
     \trestrict{r}{\rn} & \updPcAddr{\Phi\updReg{r}{c}} & If $\Phi(r) = ((\perm,\lin),\baddr, \eaddr,\aaddr)$ and $\Phi(\rn) = n$ and $\decPerm{n} \le \perm$ and $c = ((\decPerm{n},\lin),\baddr, \eaddr, \aaddr)$\\
     \hline
-    \src{\texttt{restrict} \; r \; \rn} & \src{\updPcAddr{\Phi\updReg{r}{c}}} & \srcalt{If $\Phi(r) = \stkptr{\perm,\baddr, \eaddr,\aaddr}$ and $\Phi(\rn) = n$ and $\decPerm{n} \le \perm$ and $c = \stkptr{\decPerm{n},lin),\baddr, \eaddr, \aaddr}$}\\
+   %TODO change to seal
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_3)}$\\
     \hline
+   % TODO seal
+    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
+    \hline
+
   \end{tabular}
@@
 
+\section{\srccm{} instruction interpretation}
+In this section, we present the interpretation of the \srccm{} instructions left out of \figurename~\ref{fig:target-op-sem}.
+For the instructions that only change slightly on \srccm{} compared to \trgcm{}, we include the \srccm{} specific things in blue and the things both have in common in black.
+
+ \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.2cm}|}
+    \hline
+    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
+    \hline
+    \store{r_1}{r_2}                             & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ & \update{\mem.\aaddr}{\Phi(r_2)})\endarray & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ \src{and $\aaddr \in \dom(\Phi.\mem)$}\\
+    \hline
+    \load{r_1}{r_2}                              & \left.\sourcecolor\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{w_1}\\ &\update{\mem.\aaddr}{w_\aaddr})\endarray \right.
+& \srcalt{$\Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $\aaddr \in \dom(\Phi.\ms_\stk)$ and $w_1 = \Phi.\ms_\stk(\aaddr)$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$ and $w_a = \linCons{w_1}$}\\
+    \hline
+   \geta{r_1}{r_2}                                        & \src{\updPcAddr{\Phi\updReg{r_1}{w}}} & \srcalt{If $\Phi(r_2) = ((\_,\_),\_,\_,\aaddr)$, then $w = \aaddr$ and otherwise $w = -1$}\\
+    \hline
+   \tgetb{r_1}{r_2}                                        & \src{\updPcAddr{\Phi\updReg{r_1}{w}}} & \srcalt{If $\Phi(r_2) = ((\_,\_),\baddr,\_,\_)$, then $w = \baddr$ and otherwise $w = -1$}\\
+    \hline
+   \tgete{r_1}{r_2}                                        & \src{\updPcAddr{\Phi\updReg{r_1}{w}}} & \srcalt{If $\Phi(r_2) = ((\_,\_),\_,\eaddr,\_)$, then $w = \eaddr$ and otherwise $w = -1$}\\
+    \hline
+   \tgetp{r_1}{r_2} & \src{\updPcAddr{\Phi\updReg{r_1}{w}}} & \srcalt{If $\Phi(r_2) = ((\perm,\_),\_,\_,\_)$, then $w = \encPerm{\perm}$ and otherwise $w = -1$}\\
+   \hline
+   \tsetatob{r} & \src{\updPcAddr{\Phi\updReg{r}{c}}} &\srcalt{$r \neq \pcreg$, $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\_}$, and $c = \stkptr{\perm,\baddr,\eaddr,\baddr}$} \\
+    \hline
+   \trestrict{r}{\rn} & \src{\updPcAddr{\Phi\updReg{r}{c}}} & \srcalt{If $\Phi(r) = \stkptr{\perm,\baddr, \eaddr,\aaddr}$ and $\Phi(\rn) = n$ and $\decPerm{n} \le \perm$ and $c = \stkptr{\decPerm{n},lin),\baddr, \eaddr, \aaddr}$}\\
+   \hline
+   %TODO stk
+    \splice{r_1}{r_2}{r_3}                       & \left.
+\sourcecolor\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray\right.& \srcalt{$\Phi(r_2) = \stkptr{\perm,\baddr,n,\_}$ and $\Phi(r_3) = \stkptr{\perm,n+1,\eaddr,\aaddr}$ and $\baddr \le n < \eaddr$ and $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$}\\
+    \hline
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \sourcecolor\left.\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{0} \\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray\right.
+& \srcalt{$\Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = \stkptr{\perm,\baddr,n,\aaddr}$ and $c_2 = \stkptr{\perm,n+1,\eaddr,\aaddr}$}\\
+    \hline
+
+  \end{tabular}
+
 %%% Local Variables:

### 2019-05-07 popl-2019-journal-v/appendix.tex
[POPLjv] Add world things

@@
 
+  \section{World definitions}
+  \begin{definition}
+    Given $W_{\mathrm{free}} \in \Worldfs$
+    \[
+\erase{W_{\mathrm{free}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{W_{\mathrm{free}}(r)} \in S\\
+      \bot
+    \end{array}
+  \right.
+\]
+
+  \end{definition}
+
+  \begin{definition}[Private stack sub-world erasure]
+    Given $W_{\mathrm{priv}} \in \Worlds$
+    \[
+\erase{W_{\mathrm{priv}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{\pregion{W_{\mathrm{priv}}(r)}} \in S \\
+      \bot
+    \end{array}
+  \right.
+\]
+  \end{definition}
+  \begin{definition}[Heap sub-world erasure]
+    Given $W_{\mathrm{heap}} \in \Worldh$
+    \[
+      \erase{W_{\mathrm{heap}}}{S} = \lambda r \ldotp \left\{
+        \begin{array}{l}
+          \prv{W_{\mathrm{heap}}(r)} \in S\\
+          \bot
+        \end{array}
+      \right.
+    \]
+  \end{definition}
+  \begin{definition}[World erasure]
+    Given world $(W_{\mathrm{heap}},W_{\mathrm{priv}},W_{\mathrm{free}})$ define world erasure as
+\[
+\erase{(W_{\mathrm{heap}},W_{\mathrm{priv}},W_{\mathrm{free}})}{S} =
+  \left( \erase{W_{\mathrm{heap}}}{S},\erase{W_{\mathrm{priv}}}{S},\erase{W_{\mathrm{free}}}{S} \right)
+\]
+  \end{definition}
+
+\begin{definition}[The pure part of a world]
+  \label{def:purePart}
+  For any world $W$, we define
+  \begin{align*}
+    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
+    \purePart{W_\var{heap}} &\defeq
+                       \begin{cases}
+                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_\var{heap}(r) = \revoked\\
+                       \end{cases}\\
+    \purePart{W_\var{priv}} &\defeq
+                       \begin{cases}
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\opc)\\
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\opc)\\
+                         (\revoked,\opc) & \text{if } W_\var{priv}(r) = (\revoked,\opc)\\
+                       \end{cases}\\
+    \purePart{W_\var{free}} &\defeq
+                       \begin{cases}
+                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_{\var{free}}(r) = \revoked\\
+                       \end{cases}
+  \end{align*}
+\end{definition}
+
 %%% Local Variables:

### 2019-05-07 popl-2019-journal-v/paper.tex
[POPLjv] Add oLCM missing interpretations to appendix
[POPLjv] Add world things

@@
 \renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
-\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
+\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlr{F}(#3)}
 \renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
@@
 \newcommand{\extend}[1]{}
-
+\newcommand{\Rel}[1]{\mathrm{Rel}(#1)}
+\newcommand{\fers}[1][n]{\left(\nequal[#1]\right)_{#1=0}^{\infty}}
+\newcommand{\cofe}[2][n]{\left(#2,\fers[#1] \right)}
+\newcommand{\seq}[2][n]{\left(#2_{#1}\right)_{#1=0}^{\infty}}
 \setlength{\belowcaptionskip}{-5pt}
@@
     \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
-    \hline
-    \tsplit{r_1}{r_2}{r_3}{\rn}                  & $\sourcecolor\left.\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{0} \\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray\right.$ & \srcalt{$\Phi(r_3) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = \stkptr{\perm,\baddr,n,\aaddr}$ and $c_2 = \stkptr{\perm,n+1,\eaddr,\aaddr}$}\\
     \hline
@@
 \label{sec:fa-proof}
-%\todo[inline]{Write small introduction.
-Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
-% \begin{itemize}
-% \item Logical relation
-% \item FTLR
-% \item Sketch high-level structure of the proof
-% \end{itemize}
-To prove Theorem~\ref{thm:full-abstraction}, we will essentially show that trusted components in \srccm{} are related in a certain way to their embeddings in \trgcm{}, and that untrusted \trgcm{} components are similarly related to their embeddings in \srccm{}.
-We will then prove that these relations imply that the combined programs have the same observable behavior, i.e.\ one terminates iff the other does.
-The hard part is in defining when components are related.
-In the next section, we give an informal overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
-
-\subsection{Kripke Worlds}
-\label{subsec:worlds}
-The relation between \srccm{} and \trgcm{} components is non-trivial: essentially, we will say that components are related if invoking them with related values produces related observable behavior.
-However, values are often only related under certain assumptions about the rest of the system.
-For example, the linear data part of a return capability should only be related to the corresponding \srccm{} capability if no other value in the system references the same inactive stack frame and if it is sealed with a seal that is only used for return pointers to the same code location.
-To accomodate such conditional relatedness, we construct our relation as a step-indexed Kripke logical relation with recursive worlds.
-Space constraints prevent us from explaining this in full detail in this paper. 
-Instead, we will only highlight specific important parts of the logical relation in this section, and we refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
-Additionally, for presentation, we will omit details, like step indexing, that are important for correctness but otherwise uninteresting.
-
-Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds. 
-We use a type of worlds tailored to our purposes.
-They consist of three sub-worlds: $\Wor = \Worldh \times \Worlds \times \Worldfs$, capturing assumptions about the heap\footnote{Actually, we use the term heap to describe all memory except the stack, including, for example, code memory.}, the inactive and the active part of the stack, respectively.
-Sub-worlds consist of a finite mapping from region names to regions which come in two forms (spatial and shared):
-\begin{align*}
-  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
-  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
-  \Worldfs &={} \RegionName \parfun \Regions
-\end{align*}
-Different parts of the world can contain different types of regions: heap-related assumptions can be either spatial or shared (see below), while stack-related regions must be spatial.
-Additionally, regions for inactive parts of the stack additionally include an address specifying the return address for that stack frame.
-
-Regions in $\Regionh$ specify the presence of an invariant in the system, shared with the rest of the system.
-They are tagged with the syntactic token $\pure$ and may prescribe two different types of requirements:
-\begin{align*}
-  \Regionh &= 
-             \left\{\begin{multlined}
-               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
-               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
-             \end{multlined}\right.
-\end{align*}
-First, they may require the presence of \srccm{} and \trgcm{} memory segments satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$ (readers unfamiliar with Kripke and step-indexed logical relations may read $\Wor \monnefun \URel{\MemSeg^2}$ as the set of functions from $\Wor$ to relations on memory segments).
-A region might, for example, require the presence of a certain list of instructions at a certain set of memory addresses in both \srccm{} and \trgcm{}.
-If a memory segment is owned by a given region, then that memory segment must be disjoint from the memory segments owned by all other regions.
-Second, a shared region may also contain a partial function from seals to relations on pairs of sealable capabilities: $\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps}$.
-When the region defines such a relation for a given seal, then no other region in the world can do the same, and any value signed with that seal will be required to satisfy the registered relation.
-
-Spatial regions are similar to shared regions, but they are tagged as $\spatial$ or $\spatialo$ and may not specify seal invariants.
-Additionally, a spatial region may also be $\revoked$:
-\begin{align*}
-  \Regions &={}
-             \begin{multlined}
-               \{\spatial,\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \{\revoked\}
-             \end{multlined}
-\end{align*}
-
-The difference between spatial and shared regions is related to linearity and ownership.
-For example, a \trgcm{} linear capability to a piece of memory is related to its \srccm{} counterpart, but only if no other linear capability overlaps with it.
-We will model such an assumption of exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region that only one value in the system may rely on.
-More concretely, we will define how to combine worlds $W_1$ and $W_2$ into a combined world $W_1 \oplus W_2$, on the condition that they represent compatible assumptions: $W_1$ and $W_2$ must contain the same regions except that they must respect exclusive ownership: $\pure$ regions must be present in both worlds, but a $\spatialo$ region can only be present in one and must be $\spatial$ in the other.
-When $W_1 \oplus W_2$ is defined, we say that worlds $W_1$ and $W_2$ are compatible and we refer to $W_1$ and $W_2$ as compatible partitions of the combined world.
-
-The reason that we have $\spatial$ regions (in addition to $\spatialo$ ones) is for defining when \srccm{} and \trgcm{} memories are related.
-We need them to contain suitable memory segments for all regions in the system, even for regions owned by values that live outside the memory (for example register values).
-Those regions will be in the world for the memory, but only as $\spatial$, i.e.\ they may not actually be referenced from within memory, but we still require them to be backed by suitable memory contents.
-We also use a relation $W_2 \future W_1$ that defines when a world $W_2$ is a future world of a world $W_1$.
-Relations that hold with respect to $W_1$, will then generally continue to hold in $W_2$.
-Our future world relation is fairly standard: the future world must contain all the previous world's regions, except that $\spatial$ regions are allowed to become $\spatialo$ (i.e.\ gaining ownership of a region will never break relatedness) or $\revoked$ (i.e.\ revoking spatial regions will never break relatedness).
-
-Attentive readers may have noticed that our definition of worlds is actually cyclic: worlds in $\Wor$ contain regions in $\Regionh$, but those contain partial functions from the set $\Wor$ to something else.
-Such recursive worlds are in fact common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} (essentially an advanced form of step-indexing) to construct the set $\Wor$ and rigorously resolve the circularity.
-% \begin{lemma}
-%   \label{thm:recursive-domain-eq}
-%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
-%   \[
-%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
-%   \]
-%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
-% \end{lemma}
-
-In our proof, we use only a few different types of regions.
-For space reasons, we do not go into their definitions (see the \citet{technical_report}), but we give a brief overview here.
-First, we have the code region $\codereg{\sigrets,\sigcloss,\mscode}$.
-This is a shared region that represents the assumption that memory segment
-$\mscode$ is loaded in heap memory at a certain location, it is well-formed and it uses return seals $\sigrets$ and closure seals $\sigcloss$.
-% Lau: I cannot parse this sentence:
-It takes ownership of those seals and registers appropriate invariants on the capabilities that may be signed with them.
-
-A second type of regions $\stdreg{A}{\pure}$ and $\stdreg{A}{\spatialo}$ governs heap or stack memory at a set of addresses $A$.
-It simply requires the presence of a memory segment for those addresses, such that the memory contains values that themselves satisfy the relation between values that we will see below.
-The contents of the memory is allowed to change as long as the new contents are still valid.
-Finally, a third type of regions $\stareg[(\ms_S,\ms_T)]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
-This region is used, for example, to govern inactive parts of the stack whose contents is required to remain unmodified.
-
-\subsection{The Logical Relation}
-Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
-The most important relations are summarised in the following table, where we mention the general form of the relations, what type of things they relate and extra conditions that some of them imply:\\
-% latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
-% Lau: I find it difficult to easily see what constitutes a row in this table.
-% I have updated it to a table I find easier to read.
-\newcolumntype{R}{>{$}r<{$}}
-\newcolumntype{L}{>{$}l<{$}}
-\newcolumntype{M}{R@{${}\in{}$}L}
-\begin{tabular}{|M|c|p{4.8cm}|}
-  \hline
-  \multicolumn{2}{|c|}{General form} & Relates ... & and ...\\
-  \hline
-  \npair{(w_S,w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
-  \npair{(w_S,w_T)} & \lrvtrusted(W) & values (machine words) & \\
-  \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
-  \npair{(\reg_S,\reg_T)}  &  \lrrtrusted(W) & register files & \\
-  \npair{\Phi_S,\Phi_T}  &  \lro & execution configurations & \\
-  \npair{(w_S,w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
-  \left(\arraycolsep=1pt\array{l}(w_{S,1},w_{S,2}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
-  \multicolumn{2}{|c|}{$\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
-  \hline
-\end{tabular}\\
-These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another use of step-indexing.
-For space reasons, we cannot show all of these definitions, but we will try to give an overview.
-
-Note first how we have two value relations, whose definitions are sketched in Figure~\ref{fig:value-relation}.
-The difference is that the untrusted value relation $\lrv(W)$ does not just express that the two values are related, but also that they are safe to pass to an untrusted adversary, i.e.
-they cannot be used to break LSE and WBCF.
-The trusted value relation does not have the latter requirement and is a superset of the former.
-
-Both relations trivially include numbers $(i,i)$ which are always related to themselves.
-The untrusted value relation also includes stack pointers and the underlying linear capability (with the same (non-executable) permission, range of authority, and current address), as well as syntactically equal memory capabilities, seals and sealed values, all under certain conditions involving the world $W$ and the capability's properties.
-
-Details are in the \cite{technical_report}, but roughly, for stack capabilities, the omitted condition requires that the world contains a $\spatialo$ region governing this part of the stack.
-For memory capabilities $((\perm,\lin),\baddr,\eaddr,\aaddr)$, a region in the world must govern memory $[\baddr,\eaddr]$, either $\spatialo$ or $\pure$, depending on the linearity $\lin$ of the capability.
-If the capability is executable ($\perm \in \{\rx,\rwx\}$), then we additionally require that the governing region is a code region and that the two capabilities are related $\mathrm{jmp}$ targets, as expressed by the relation $\lre(W)$, in any future world (see below).
-
-Seals allocated to trusted code are related to themselves only by $\lrvtrusted(W)$, but other seals are in both value relations.
-Sealed values are in both relations essentially when the sealed values satisfy the relation that was registered for the seal in a region of the world.
-Additionally, when they are combined with any other pair of values related by that relation, they must be related as $\txjmp{}{}$ targets (i.e. in $\lrexj(W)$).
-Finally, capabilities to code memory are related to themselves in the trusted value relation ($\lrvtrusted(W)$) when there is an appropriate code region in the world.
-They are not in the untrusted value relation because the code memory contains copies of the return seals used by the code, which must not end up in the hands of an adversary.
-
-\begin{figure}
-  \centering
-  \begin{align*}
-  \lrv(W) ={} & \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\ &
-%
-    \left\{
-%    \begin{array}{l}
-      \npair{\left(\src{\stkptr{\perm,\baddr,\eaddr,\aaddr}}, ((\perm,\linear),\baddr,\eaddr,\aaddr) \right)} \mid\dots
-        % \perm \not\in \{\rx,\rwx\} \tand \\
-        % % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in
-        % % \lrp(W) \wedge \\
-        % \quad\perm \in \{\ro,\rw\} \Rightarrow \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \tand \\
-        % \quad\perm = \rw  \Rightarrow \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
-%    \end{array}
-    \right\} \cup \\ &
-%
-    \left\{
-%    \begin{array}{l}
-      \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} \mid 
-      % [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss) \wedge 
-                       \dots 
-      % \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      % \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
-%    \end{array}
-    \right\} \cup \\ &
-        \left\{
-%    \begin{array}{l}
-      \npair{\left(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T} \right)} \mid \dots
-      % \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
-      % \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      % \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      % \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
-      % \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      % \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      % \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \tand \dots
-      % % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      % % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      % % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-%    \end{array}
-    \right\}\cup\\ &
-%
-     \left\{ \npair{\left(\src{((\perm,\lin),\baddr,\eaddr,\aaddr)}, ((\perm,\lin),\baddr,\eaddr,\aaddr)\right)} \mid \dots
-    % \begin{array}{l}
-    %   [b,e] \mathrel{\#} \ta \tand\\
-    %   \begin{array}{r l l }
-    %     % \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in
-    %     % \lrp(W) \wedge \\
-    %     \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
-    %     \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
-    %     % we are excluding rwx pointers.
-    %     \perm \neq \rwx \wedge \\
-    %     % \perm = \rwx &\Rightarrow&
-    %     % \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W}
-    %     % \wedge \\
-    %     % \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
-    %     \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
-    %     \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
-    %     \lin = \normal \\ \endarray
-    %   \end{array}
-    % \end{array}
-     \right\} \\
-%  \end{array}
-  \lrvtrusted(W) ={} & \lrv(W)\cup \\
-%  \begin{array}[t]{l}
-    &\left\{
-    %\begin{array}{l}
-      \npair{\left(\src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}}, \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \right)} \mid
-      % [\sigma_\baddr,\sigma_\eaddr] \subseteq(\sigrets \cup \sigcloss) \wedge 
-      \dots 
-    %           \exists r \in \dom(\pwheap) \ldotp \\
-    %           \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\mscode,\gc} \tand \dom(\mscode) \subseteq \ta \\
-    %           \quad \tand [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \tand \sigrets \subseteq \gsigrets \tand \sigcloss \subseteq \gsigcloss
-    % \end{array}
-    \right\} \cup \\
-    & \left\{
-%    \begin{array}{l}
-      \npair{\left(\src{((\perm,\normal),\baddr,\eaddr,\aaddr)},((\perm,\normal),\baddr,\eaddr,\aaddr) \right)} \mid \perm \le \rx \wedge \dots
-    %   \quad \perm \sqsubseteq \rx \tand 
-    %    [\baddr,\eaddr] \subseteq \ta \tand 
-    %    \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
-    % \end{array}
-    \right\}
-%  \end{array}
-\end{align*}
-\caption{Sketches of the trusted and untrusted value relation.}
-\label{fig:value-relation}
-\end{figure}
-
-\begin{figure}
-  \centering
-\begin{align*}
-  \lrrg{\trust}(W) &= \left\{ \npair{\stpair{\reg}{\reg}} \middle|
-    \begin{array}{l}
-    % Lau: Consider using \Wor instead of \World as we have not made a distinction between the two.
-      \exists S : (\RegName \setminus \{\pcreg \})\fun \World \ldotp \\
-      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg \} \cup R))} S(r) \wedge \\
-      \quad \forall r \in \RegName \setminus \{\pcreg \}\ldotp \npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
-    \end{array}
-            \right\}\\
-  \lre(W)&= \left\{ \begin{array}{l}
-    \npair{\stpair[.]{w_{c,S}}{w_{c,T}}} | \\
-    \quad\forall \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-    \qquad\npair{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
-    \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{w_{c,S}} \tand\\
-                     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{w_{c,T}} \tand \\
-                     \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-    \qquad\qquad\Rightarrow\npair{\left(\Phi_S', \Phi_T' \right)}\in \lro
-  \end{array}
-  \right\}
-\end{align*}
-  \begin{align*}
-  \lro[] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
-    \src{\Phi_S \sterm{}} \Leftrightarrow \Phi_T \trg{\term} \}
-\end{align*}
-\caption{Simplified sketches of the register file relation $\lrr(W)$, the relation
-  for $\com{jmp}$ targets $\lre(W)$  and the observation relation $\lro(W)$.}
-\label{fig:obs-rel}
-\end{figure}
-
-In Figure~\ref{fig:obs-rel}, we show sketches of the register file relation $\lrr$, the relation between $\tjmp{}$ targets $\lre(W)$ and the observation relation $\lro$.
-The trusted/untrusted register file relation simply requires that all registers except $\pcreg$ are in the corresponding value relation (in a compatible world partition).
-Two execution configurations are in the observation relation $\lro$ if one terminates whenever the other does\footnote{The actual definition in the \cite{technical_report} is complicated a bit by step-indexing and the fact that we actually use two separate observation relations for left- and right-approximation.}.
-The $\lre(W)$ relation then includes any two words which can be plugged into related register files and memories (in compatible worlds), to obtain execution configurations in the observation relation.
-
-\subsection{Fundamental Theorem}
-An important lemma in our proof of full abstraction of the embedding of \srccm{} into \trgcm{}, is the fundamental theorem of logical relations (FTLR).
-The name indicates that it is an instance of a general pattern in logical relations proofs, but is otherwise unimportant.
-\begin{lemma}[FTLR (roughly)]
-  \label{thm:ftlr}
-  If $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and $\pwheap(r) = \codereg{\sigrets,\sigcloss,\mscode}$, and either $[\baddr,\eaddr] \subseteq \ta$ and $\mscode$ behaves reasonably (see Section~\ref{sec:well-form-reas}) or
-$[\baddr,\eaddr] \mathrel{\#} \ta$,
-then 
-  \[
-    \npair{\left(\src{((\rx,\normal),\baddr,\eaddr,\aaddr)}, ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
-  \]
-\end{lemma}
-
-Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
-The conditions require that the capability points to a memory region where code is loaded and that code must be either trusted and behave reasonably (i.e.\ respect the restrictions that \stktokens{} relies on, see Section~\ref{sec:well-form-reas}) or untrusted (in which case, it cannot have WBCF or LSE expectations, see Section~\ref{sec:well-form-reas}).
-
-The proof of the lemma consists of a big induction where each possible instruction is proven to behave the same in source and target in related memories and register files.
-After that first step, the induction hypothesis is used for the rest of the execution.
-
-\subsection{Full Abstraction Proof Sketch}
-\label{subsec:proof-sketch}
-Using Lemma~\ref{thm:ftlr}, we can now proceed to proving Theorem~\ref{thm:full-abstraction} (full abstraction).
-First, we extend the logical relation into an omitted relation on components $(\var{comp}_S,\var{comp}_T) \in \lrcomp(W)$.
-% Essentially, it relates a component to itself if instantiating their imports with related values produces related exports and code memory satisfying the appropriate code region in the world.
-% The component relation $\mathcal{C}(W)$ basically lifts the logical relation we have presented above to components.
-% The component relation relates a component $(\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss)$ to itself when two conditions are satisfied.
-% First, when words that relate to them selves in the untrusted value relation are used to satisfy the imports, i.e.\ the words are placed on the import addresses in $\msdata$, and this data memory is combined with the code memory $\mscode$, then it forms a safe heap, i.e.\ it is in the $\mathcal{H}$ relation.
-% Second, the exports should always be safe to use which means that they must be in the untrusted value relation in any future world.
-% \begin{multline*}
-%   \lrcomp(W) =\\
-%   \left\{\begin{aligned}
-%       &\npair{\var{comp},\var{comp}} \;\mid \;\\
-%       &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
-%       &\qquad\text{For all } W' \future W \ldotp \\
-%       &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
-%       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-%       &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
-%       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
-%     \end{aligned}
-%   \right\}\\
-% \cup \left\{
-%     \begin{multlined}
-%       \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
-%       \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand
-%       \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
-%     \end{multlined}
-%   \right\} 
-% \end{multline*}
-%
-Using Lemma~\ref{thm:ftlr} and the definitions of the logical relations, we can then prove the following two lemmas.
-The first is a version of the FTLR for components, stating that all components are related to themselves if they are either (1) well-formed and untrusted or (2) well-formed, reasonable and trusted.
-\begin{lemma}[FTLR for components]
-  \label{lem:ftlr-comps}
-  If $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$ and either
-    $\dom(\comp.\mscode) \subseteq \ta$ and $\src{\comp}$ is a reasonable component; or
-    $\dom(\comp.\mscode) \mathrel{\#} \ta$,
-  then there exists a $W$ such that
-  $\npair{(\src{\comp},\comp)} \in \lrcomp(W)$.
-\end{lemma}
-
-Another lemma then relates the component relation and context plugging: plugging related components into related contexts produces related execution configurations.
-\begin{lemma}
-  \label{lem:adeq-context-plug}
-  If $\npair{\stpair{\context}{\context}}\in\lrcomp(W_1)$ and $\npair{\stpair{\comp}{\comp}} \in \lrcomp(W_2)$ and $W_1\oplus W_2$ is defined, then
-  $\plug{\src{\context_S}}{\src{\comp_S}}$ terminates iff $\plug{\context_T}{\comp_T}$ terminates.
-\end{lemma}
-% Finally, we have an adequacy lemma for the execution configuration relation.
-% This lemma says that for related \srccm{} and \trgcm{} configurations, one terminates iff the other does.
-% \begin{multline*}
-%   \lrec[\square,\gc = (\ta,\stkb)](W) = \\
-% \left\{
-%   \begin{array}{l}
-%      \npair{\left(
-%     (\ms_S,\reg_S,\stk,\ms_\stk),
-%     (\ms_T,\reg_T)\right)} \mid \\
-%     \quad \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
-%     \qquad \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
-%     \qquad \reg_S(\pcreg) \neq \retptrc(\_) \tand
-%     \reg_S(\rdata) \neq \retptrd(\_) \tand \\
-%     \qquad \nonExec{\reg_S(\rdata)} \tand  
-%      \nonExec{\reg_T(\rdata)} \tand \\
-%     \qquad \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
-%   \end{array}
-% \right\}
-% \end{multline*}
-% \begin{lemma}[Adequacy of execution configuration logical relation]
-%   \label{lem:adequacy}
-%   If $\npair{\stpair{\Phi}{\Phi}}\in\lrec(W)$ then $\src{\Phi_S} \sterm{}$ iff $\Phi_T\term$.
-% \end{lemma}
-
-Finally, we use these two lemmas to prove Theorem~\ref{thm:full-abstraction}.
-\begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  The proofs of both directions are similar, so we only show the right direction.
-  To show the \trgcm{} contextual equivalence, assume w.l.o.g\ a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.
-  The proof is sketched in Figure~\ref{fig:fa-proof-sketch}.
-  By the statement of Theorem~\ref{thm:full-abstraction}, we may assume that the trusted components $\src{\comp_1}$ and $\src{\comp_2}$ are well-formed and reasonable.
-  We prove arrow (1) in the figure by using the mentioned assumptions about $\src{\comp_1}$ and $\trg{\context}$ along with Lemma~\ref{lem:ftlr-comps} and \ref{lem:adeq-context-plug}.
-  Now we know that ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{}{}}$, so by the assumption that $\src{\comp_1}$ and $\src{\comp_2}$ are contextually equivalent on \srccm{} we get ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{}{}}$, i.e.\ arrow (2) in the figure.
-  To prove arrow (3), we again apply Lemma~\ref{lem:ftlr-comps}, \ref{lem:adeq-context-plug}; but this time, we use the assumption that $\src{\comp_2}$ is well-formed and reasonable and that $\trg{\context}$ is well-formed.
-\end{proof}
-
-\renewcommand{\comp}{C}
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
-    % \draw[help lines,yellow] (0,0) grid (10,7);
-    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };
-
-    \node at (3.4,4) { ${\plug{\trg{\context}}{\src{\comp_1}} \sterm[]{\gc}{}}$ };
-    \node at (5,4) { $\mathrel{\Rightarrow}$ };
-    \node at (6.6,4) { ${\plug{\trg{\context}}{\src{\comp_2}} \sterm[]{\gc}{}}$ };
-
-    \node at (4.35,2.8) { (1) };
-    \node at (5,3.6) { (2) };
-    \node at (5.65,2.8) { (3) };
-
-    \draw[out=100,in=260,double,-implies,double equal sign distance] (4,2.6) to (4,3.4);
-
-    \draw[out=280,in=80,double,-implies,double equal sign distance] (6,3.4) to (6,2.6);
-
-    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
-    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
-      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
-    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
-    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
-    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };
-
-    \node at (3.4,2) { ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$ };
-    \node at (5,2.1) { $\overset{?}{\Rightarrow}$ };
-    \node at (6.6,2) { ${\plug{\trg{\context}}{\src{\comp_2}} \term[]{}}$ };
-
-    \node at (5,1.3) { ${\src{\comp_1}}\mathrel{\overset{?}{\tconeq}}{\src{\comp_2}}$ };
-
-    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\Small Contextual equivalence preservation} (0,1);
-  \end{tikzpicture}
-  \caption{Proving one direction of fully abstract compilation (contextual equivalence preservation).}
-  \label{fig:fa-proof-sketch}
-\end{figure}
-
-% \subsection{Proof sketch}
-% \label{subsec:proof-sketch}
-% \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  % \item Consider first the upward arrow.
-  %   Assume $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$.
-
-  %   Take a $\src{\context}$ such that $\vdash \src{\context}$, take $\src{\ta[,i]}
-  %   = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and
-  %   $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$ and we will prove that
-  %   $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}} \Leftrightarrow
-  %   \src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
-
-  %   By symmetry, we can assume w.l.o.g. that $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$ and prove that $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
-  %   Note that this implies that $\src{\context}$ is a valid context for both $\src{\var{comp}_1}$ and $\src{\var{comp}_2}$.
-
-  %   First, we show that also $\plug{\context}{\var{comp}_1} \trg{\term}$.
-  %   Take $n$ the amount of steps in the termination of $\src{\plug{\context}{\var{comp}_1} \sterm{\gc_1}}$.
-  %   It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\preceq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-  %   It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
-  %   Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\preceq,\gc_1](W_1\uplus W_1')$
-  %   Together with $\src{\plug{\context}{\var{comp}_1} \sterm[n]{\gc_1}}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \trg{\term}$.
-
-  %   It follows from $\src{\var{comp}_1} \tconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
-
-  %   It now remains to show that also $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$.
-  %   Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \trg{\term}$.
-  %   It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\succeq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-  %   It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
-  %   Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\succeq,\gc_2](W_2\uplus W_2')$
-  %   Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\src{\plug{\context}{\var{comp}_2} \sterm{\gc_2}}$, concluding this direction of the proof.
-
-%   First consider the right arrow:
-
-%     Assume $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$.
-Take $\src{\ta[,i]} = \src{\dom(\var{comp}_i.\mscode)}$, $\gsigrets_i = \var{comp}_i.\sigrets$ and $\gsigcloss_i = \var{comp}_i.\sigcloss$, $\gc_i = (\ta[,i],\stkb_i,\gsigrets_i,\gsigcloss_i)$.
-% %
-%     Take a $\trg{\context}$ such that $\vdash \trg{\context}$ and we will prove that
-%     $\trg{\plug{\context}{\var{comp}_1} \term} \Leftrightarrow
-%     \trg{\plug{\context}{\var{comp}_2} \term}$.
-% %
-%     By symmetry, we can assume w.l.o.g. that $\trg{\plug{\context}{\var{comp}_1} \term}$ and prove that $\trg{\plug{\context}{\var{comp}_2} \term}$.
-%     Note that this implies that $\trg{\context}$ is a valid context for both $\trg{\var{comp}_1}$ and $\trg{\var{comp}_2}$.
-% %
-%     First, we show that also $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
-%     Take $n$ the amount of steps in the termination of $\plug{\context}{\var{comp}_1} \trg{\term}$.
-%     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\var{comp}_1,\var{comp}_1)} \in \lrcomp[\succeq,\gc_1](W_1)$ for some $W_1$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-%     It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n+1]{(\context,\context)} \in \lrcomp[\succeq,\gc_1](W_1')$ for some $W_1'$ that we can choose such that $W_1 \uplus W_1'$ is defined.
-%     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair{(\plug{\context}{\var{comp}_1}, \plug{\context}{\var{comp}_1})} \in \lrec[\succeq,\gc_1](W_1\uplus W_1')$
-%     Together with $\plug{\context}{\var{comp}_1} \trg{\term[n]}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_1} \src{\sterm{\gc_1}}$.
-% %
-%     It follows from $\src{\var{comp}_1} \sconeq \src{\var{comp}_2}$ that also $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
-% %
-%     It now remains to show that also $\plug{\context}{\var{comp}_2} \trg{\term}$.
-%     Take $n'$ the amount of steps in the termination of $\plug{\context}{\var{comp}_2} \src{\sterm{\gc_2}}$.
-%     It follows from Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\var{comp}_2,\var{comp}_2)} \in \lrcomp[\preceq,\gc_2](W_2)$ for some $W_2$ with $\dom(\pwfree) = \dom(\pwpriv) = \emptyset$.
-%     It also follows from the same Lemma~\ref{lem:ftlr-comps} that $\npair[n'+1]{(\context,\context)} \in \lrcomp[\preceq,\gc_2](W_2')$ for some $W_2'$ that we can choose such that $W_2 \uplus W_2'$ is defined.
-%     Lemma~\ref{lem:compat-context-plug} then tells us that $\npair[n']{(\plug{\context}{\var{comp}_2}, \plug{\context}{\var{comp}_2})} \in \lrec[\preceq,\gc_2](W_2\uplus W_2')$
-%     Together with $\plug{\context}{\var{comp}_2} \trg{\term[n']}$, Lemma~\ref{lem:adequacy} then tells us that $\plug{\context}{\var{comp}_2} \trg{\term}$, concluding the second direction of the proof.
-
-%  The left arrow is proven in a similar manner.
-% \end{proof}
+\input{proving-full-abstraction}
 

### 2019-05-07 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add world things

--- 2018-11-09 fully-abstract-comp/popl2019/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,screen]{acmart}\settopmatter{}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986.10002989</concept_id>
-<concept_desc>Security and privacy~Formal security models</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10003006</concept_id>
-<concept_desc>Security and privacy~Systems security</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10010124.10010138</concept_id>
-<concept_desc>Theory of computation~Program reasoning</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Security and privacy~Formal security models}
-\ccsdesc[300]{Security and privacy~Systems security}
-\ccsdesc[300]{Theory of computation~Program reasoning}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-% \acmJournal{PACMPL}
-% \acmVolume{1}
-% \acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
-% \acmArticle{1}
-% \acmYear{2018}
-% \acmMonth{1}
-% \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-% \startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-% \setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-\usepackage[utf8]{inputenc}
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-% %% Some recommended packages.
-% \usepackage{booktabs}   %% For formal tables:
-%                         %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-\usepackage{wrapfig}
-
-\usepackage{placeins}
-
-\usepackage[disable]{todonotes}
-
-\input{preamble-illu}
-
-\input{../preamble}
-% Paper specific redefinitions of commands
-\renewcommand{\npair}[2][n]{#2}
-\renewcommand{\MemSeg}{\shareddom{MemSeg}}
-\renewcommand{\Reg}{\shareddom{RegFile}}
-\renewcommand{\RegName}{\shareddom{RegName}}
-\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
-\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
-\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
-\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
-\renewcommand{\perm}{\var{p}}
-\renewcommand{\SealableCaps}{\shareddom{Sealables}}
-\renewcommand{\Cap}{\shareddom{NatTok}}
-\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
-\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
-\renewcommand{\comp}{\var{comp}}
-\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
-\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
-\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
-\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
-
-\renewcommand{\lre}[1][]{\typesetlr{E}^{#1}}
-\renewcommand{\lretrusted}[1][]{\lreg[#1]{\trusted}}
-\renewcommand{\lrexj}[1][]{\typesetlr{E}^{#1}_{\mathrm{xjmp}}}
-\renewcommand{\lrk}[1][]{\typesetlr{K}^{#1}}
-\renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
-\renewcommand{\lrr}[1][]{\lrrg[#1]{\untrusted}}
-\renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
-\renewcommand{\lro}[1][]{\typesetlr{O}^{#1}}
-\renewcommand{\lrol}{\lro[\preceq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
-\renewcommand{\lror}{\lro[\succeq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
-\renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
-\renewcommand{\lrv}[1][]{\lrvg[#1]{\untrusted}}
-\renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
-\renewcommand{\lrp}{\typesetlr{P}}
-
-\renewcommand{\lrrs}{\typesetlr{R}}
-\renewcommand{\lrm}{\typesetlr{M}}
-\renewcommand{\lrcomp}[1][]{\typesetlr{C}^{#1}}
-\renewcommand{\lrec}[1][]{\typesetlr{EC}^{#1}}
-\renewcommand{\lrheap}[1][]{\typesetlr{H}^{#1}}
-
-\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
-\renewcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
-\renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
-\renewcommand{\codereg}[2][\mathrm{code}]{\iota^{#1}_{#2}}
-\renewcommand{\staureg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta,\lrv},#2}_{#1}}
-
-\newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
-\newcommand{\srcxjmpres}[1]{\plainfun{\srcalt{xjmpRes}}{#1}}
-\newcommand{\wdjud}[2][ ]{#1 \vdash #2}
-
-\newcommand{\trgcm}{\textsc{LCM}}
-\newcommand{\srccm}{\textsc{oLCM}}
-\newcommand{\extend}[1]{}
-
-\setlength{\belowcaptionskip}{-5pt}
-
-%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
-%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
-%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
-%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
-%%%
-%%% The following is specific to POPL '19 and the paper
-%%% 'StkTokens: Enforcing Well-Bracketed Control Flow and Stack Encapsulation using Linear Capabilities'
-%%% by Lau Skorstengaard, Dominique Devriese, and Lars Birkedal.
-%%%
-\setcopyright{rightsretained}
-\copyrightyear{2019}
-\acmJournal{PACMPL}
-\acmYear{2019}
-\acmVolume{3}
-\acmNumber{POPL}
-\acmArticle{19}
-\acmMonth{1}
-\acmPrice{}
-\acmDOI{10.1145/3290332}
-
-\begin{document}
-
-%% Title information
-\title[\stktokens{}: Enforcing WBCF and LSE using linear capabilities]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
-% \titlenote{with title note}
-% \subtitle{Fully abstract overlay semantics}
-% \subtitlenote{with subtitle note}
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-\author{Lau Skorstengaard}
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \institution{Aarhus~University}
-  \country{Denmark}                    %% \country is recommended
-}
-\email{lask@cs.au.dk}
-
-%% Author with two affiliations and emails.
-\author{Dominique Devriese}
-\orcid{0000-0002-3862-6856}             %% \orcid is optional
-\affiliation{
-  \institution{Vrije~Universiteit~Brussel}           %% \institution is required
-  \country{Belgium}                   %% \country is recommended
-}
-\email{dominique.devriese@vub.be}         %% \email is recommended
-
-\author{Lars Birkedal}
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \institution{Aarhus~University}
-  \country{Denmark}                    %% \country is recommended
-}
-\email{birkedal@cs.au.dk}
-
-\begin{abstract}
-  We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
-  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
-  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalize and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
-\end{abstract}
-
-% %% 2012 ACM Computing Classification System (CSS) concepts
-% %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-% \begin{CCSXML}
-% <ccs2012>
-% <concept>
-% <concept_id>10011007.10011006.10011008</concept_id>
-% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-% <concept_significance>500</concept_significance>
-% </concept>
-% <concept>
-% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-% <concept_significance>300</concept_significance>
-% </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Software and its engineering~General programming languages}
-% \ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, fully abstract overlay semantics}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-\todo[inline]{TODOS ENABLED!!!}
-\todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be made.
-Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
-Generally, properties that hold in the source language but not in the target language need to be somehow enforced by the compiler.
-Two properties that hold in many high-level source languages, but not in the assembly languages they are compiled to, are well-bracketed control flow and encapsulation of local state.
-
-Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
-At the assembly level, this property does not hold immediately. 
-Invoked functions get direct access to return pointers that they are supposed to jump to a single time at the end of their execution.
-There is, however, no guarantee that untrusted assembly code respects this intended usage.
-In particular, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.
-
-Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not be read or modified until the invoked function returns.
-At the assembly level, this property also does not hold immediately.
-The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
-However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
-
-To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
-The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
-More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
-These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
-Capabilities come in different flavours.
-Memory capabilities allow reading from and writing to a block of memory.
-Additionally, capability machines offer some form of {\itshape object capabilities} that represent low-level encapsulated closures, i.e.
-a piece of code coupled with private state that it gains access to upon invocation.
-The concrete mechanics of object capabilities vary between different capability machines.
-% KJAA: This 'sealed' and 'common seal' business is unclear to me.
-For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
-Each of the two capabilities are sealed with a common seal which make them opaque.
-The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.
-
-To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
-The first is to use separate stacks for mutually distrusting components, and a central, trusted stack manager that mediates cross-component invocations.
-This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it requires reserving separate stack space for all components, which scales poorly to large amounts of components.
-Also, in the presence of higher-order values (e.g., function pointers, objects), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
-It is not clear how it can do this efficiently in the presence of large amounts of components.
-Finally, this approach does not allow passing stack references between components.
-
-A more scalable approach retains a single stack shared between components.
-Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component and to revoke them after it returns.
-While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
-For example, CHERI offers a form of {\itshape local} capabilities that can only be stored in registers or on the stack but not in other parts of memory.
-\citet{skorstengaard_reasoning_2017} have demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
-In fact, a similar system was envisioned in earlier work on CHERI~\citep{watson2012cheri}. 
-% KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.
-
-In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
-Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson2018ISA}.
-% KJAA: Perhaps make it clear by who and when/where?
-The hardware prevents these capabilities from being duplicated.
-We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to return them in returns.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
-
-A second contribution of this work is the way in which we formulate these two properties.
-Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
-Existing formulations are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
-We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
-It starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To this end, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
-This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
-As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
-To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
-More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
-
-Compared to~\citet{skorstengaard_reasoning_2017} that prove LSE and WBCF for a handful of examples,
-this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
-Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
-The idea is that such a compiler could be proven fully abstract with respect to the well-behaved semantics of the target language, so that the proof could rely on native well-bracketedness and local stack frame encapsulation.
-Such an independent result could then be composed with ours to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
-
-%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
-In this paper, we make the following contributions:
-% Presents the first formal model of a capability machine with linear capabilities (Section 2)
-% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
-% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
-\begin{itemize}
-\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
-\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present a new way to formalize these guarantees based on a novel
-  technique called \textit{fully-abstract overlay semantics} and we prove LSE
-  and WBCF claims. This includes:
-  \begin{itemize}
-  \item \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
-  \item proving full-abstraction for the embedding of \srccm{} into \trgcm{} (Section~\ref{sec:fa-proof}) by
-  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
-  \end{itemize}
-\end{itemize}
-This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.
-
-%\paragraph{Outline} Blabla
-
-\section{A Capability Machine with Sealing and Linear Capabilities}
-\label{sec:cap-mach-w-seal-and-lin}
-% Brief intro to section and this cap machine:
-% Cap machine inspired by CHERI, but with linear capabilities
-In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
-For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
-% \lau{Do we have a good reason to assume an infinite address space for this work?
-%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
-
-%%% Capabilities
-The concept of a capability is the cornerstone of any capability machine.
-In its simplest form, a capability is a permission and a range of authority.
-The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
-% Permissions and Range of authority
-The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
-Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
-The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$) ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
-% Pointer and linearity
-In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
-The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
-% What is linear capabilities, split/splice
-A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
-The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
-
-\begin{wrapfigure}{r}{0.19\linewidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (rwx) {$\rwx$};
-    \node[main node] (rx) [below right of=rwx] {$\rx$};
-
-    \node[main node] (rw) [below left of=rwx,] {$\rw$};
-    \node[main node] (r) [below right of=rw] {$\readonly$};
-    \node[main node] (0) [below of=r] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-
-    (rw) edge (r)
-    (r) edge (0)
-
-    (rwx) edge (rx)
-
-    (rw) edge (rwx)
-    (r) edge (rx);
-  \end{tikzpicture}
-
-  \caption{Permission hierarchy}
-  \label{fig:perm-hier}
-\end{wrapfigure}%
-%
-%%% Seals
-% KJAA: Here I am still missing an intuition of what a seal 'is' -- why the range, why the seal as an 'entity'?
-
-Any reasonable capability machine needs a way to set up boundaries between security domains with different authorities.
-It also must have a way to cross these boundaries such that (1) the security domain we move from can encapsulate and later regain its authority and (2) the security domain we move to regains all of its authority.
-%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015,watson_fast_2016}.
-% seals
-A sealed capability makes an underlying capability opaque which means that the underlying capability cannot be changed or used for the operations it normally gives permission to.
-In other words, the authority the underlying capability represents is encapsulated under the seal.
-Syntactically, the sealed capability is represented as $\sealed{\sigma,\vsc}$ where $\vsc$ is a sealable and $\sigma$ is the seal used to seal the capability.
-In order to seal a sealable with a seal $\sigma$, it is necessary to have the right to do so.
-The right to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
-A set of seals is a capability that represents the authority to seal sealables with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In spirit of memory capabilities, a set of seals has an active seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
-% On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
-As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}, an operation that operates on a pair of capabilities sealed with the same seal.
-The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part of the pair) available to the invoked code.
-The combination of sealed capabilities and \texttt{xjmp} gives (1) and (2).
-\begin{figure}[tb]
-  \centering
-  \[
-  \arraycolsep=1.4pt
-  \begin{array}{rrcl l rrcl}
-   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
-    \aendbnf \in & & & \Addr \cup \{\infty \} & & \sealendbnf \in & & & \Seal \cup \{\infty \}\\
-    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
-    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
-    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
-    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
-    \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
-\ms \in  &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
-    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
-  \end{array}
-\]
-\[
-  \arraycolsep=1.4pt
-\begin{array}{rcl}
-\multicolumn{3}{l}{    \arraycolsep=0pt
-      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
-}\\
-  \Instr &\defbnf & \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tload{\com{r}}{\com{r}} \mid \tstore{\com{r}}{\com{r}} \mid \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid\\
-         & & \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid\tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}}  \mid \\
-  & & \tcca{\com{r}}{\com{n\rn}} \mid \tsetatob{\com{r}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tcseal{\com{r}}{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid  \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\
-      & & \tsplice{\com{r}}{\com{r}}{\com{r}} \mid \tfail \mid \thalt 
-\end{array}
-\]
-  \caption{The syntax of our capability machine with seals and linear capabilities.}
-  \label{fig:target-syntax}
-\end{figure}
-
-%%% Domains
-Words on \trgcm{} are capabilities and data (represented by integers ($\ints$)).
-We assume a finite set of register names $\RegName$ containing at least $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$, and we define register files as functions from register names to words.
-Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ partial functions).
-$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
-An executable configuration is a register file and memory pair.
-
-\begin{figure}[p]
-  \centering
-  \begin{mathpar}
-    \inference{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \\
-        \baddr \le \aaddr \le \eaddr \and \perm \in \{\rwx,\rx\}
- }{ \Phi \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) }
-              \and
-              \inference{ \forall \Phi' \neq \failed \ldotp \Phi \nrightarrow \Phi'}
-                        {\Phi \step \failed }
-  \end{mathpar}
-  \[
-    \begin{array}{l}
-  %   \Phi \step
-  % \begin{cases}
-  %   \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) &
-  %   \arraycolsep=0pt
-  %     \begin{array}[t]{l}
-  %       \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \wedge
-  %       \baddr \le \aaddr \le \eaddr \wedge \perm \in \{\rwx,\rx\}
-  %     \end{array} \\
-  %     \failed & \totherwise
-  % \end{cases}\\
-  \updPcAddr{\Phi} =
-  \begin{cases}
-    \Phi\updReg{\pcreg}{w} & \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge w = ((\perm,\lin),\baddr,\eaddr,\aaddr+1)\\
-    \Phi  & \totherwise
-  \end{cases}\\
-  \linCons{w} =
-  \begin{cases}
-    0 & \isLinear{w} \\
-    w & \totherwise
-  \end{cases}\\
-  \xjmpres{c_1,c_2,\Phi} =
-  \begin{cases}
-    \Phi\updReg{\pcreg}{c_1}\updReg{\rdata}{c_2} & \nonExec{c_2} \\
-    \failed & \totherwise
-  \end{cases}
-  \end{array}
-\]
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.7cm}|}
-    \hline
-    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
-    \hline
-    \halt                                        & \halted & \\
-    \hline
-    \fail                                        & \failed & \\
-    % \hline
-    % \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{r}{\rn}} & $\rn \in \ints$\\
-    \hline
-    \move{r}{\rn}                                & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\Phi(\rn)}$ \\
-    \hline
-    \load{r_1}{r_2}                              & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{w_1}\\ &\update{\mem.\aaddr}{w_\aaddr})\endarray & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$ and $w_a = \linCons{w_1}$\\
-    \hline
-    \store{r_1}{r_2}                             & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ & \update{\mem.\aaddr}{\Phi(r_2)})\endarray & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
-    \hline
-    \geta{r_1}{r_2}                              & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\_,\aaddr)$ or $\Phi(r_2) = \seal{\_,\_,\aaddr}$, then $w = \aaddr$ and otherwise $w = -1$\\
-    \hline
-    \cca{r}{\rn}                                 &\updPcAddr{\Phi\updReg{r}{w}} & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
-    \hline
-    \jmp{r}    &\Phi\arraycolsep=0pt\array[t]{l}\updReg{r}{w}\\\updReg{\pcreg}{\Phi(r)}\endarray & $w = \linCons{\Phi(r)}$\\
-    \hline
-    \xjmp{r_1}{r_2}                              & \Phi' & $\Phi(r_1) = \sealed{\sigma,c_1}$ and $\Phi(r_2) = \sealed{\sigma,c_2}$ and $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$ and $\Phi' = \xjmpres{c_1,c_2,\Phi\updReg{r_1,r_2}{w_1,w_2}}$  \\
-    \hline
-    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_3)}$\\
-    \hline
-    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
-    \hline
-    \cseal{r_1}{r_2}                             & \updPcAddr{\Phi\updReg{r_1}{\vsc}} & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
-    \hline
-    \multicolumn{3}{|c|}{\dots} \\
-    \hline
-    \_                                           & \failed & \totherwise \\
-    \hline
-  \end{tabular}
-\caption{An excerpt of the operational semantics of \trgcm{}}
-  \label{fig:target-op-sem}
-\end{figure}
-% syntax (instructions)
-\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machines as well as capability-related instructions.
-The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability instructions inspect the properties of capabilities: type (\texttt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
-The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
-The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
-Generally speaking, a capability machine needs an instruction for reducing the range of authority of a capability.
-Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two rather than reducing the range of authority (\texttt{split}).
-The reverse is possible using \texttt{splice}.
-Sealables can be sealed using \texttt{cseal} and pairs of sealed capabilities can be unsealed by crossing security boundaries (\texttt{xjmp}, see below).
-Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).
-
-% opsem
-The operational semantics of \trgcm{} is displayed in Figure~\ref{fig:target-op-sem}.
-The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
-The executed instruction is determined by the capability in the $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
-In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
-If both conditions are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
-The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
-In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
-The instructions that stop execution or change the flow of execution do not use $\updPcAddr{}$.
-For instance, the \texttt{halt} and \texttt{fail} instructions are simply interpreted as the $\halted$ and $\failed$ configurations, respectively, and they do not use $\updPcAddr{}$.
-
-% move w/ lin
-The \texttt{move} instruction simply moves a word from one register to another.
-It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
-When a linear capability is moved, the source register must be cleared, so the capability is not duplicated.
-This is done uniformly in the semantics using the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
-When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
-In the case of \texttt{move}, the source register $\rn$ is overwritten with $\linCons{\Phi(\rn)}$.
-
-% load w/ lin (and store)
-The \texttt{store} and \texttt{load} instructions are fairly standard.
-They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority.
-Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
-In this case, we require that the memory capability used for loading also has write-permission.
-
-% geta, cca
-The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
-\texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
-Note that this instruction does not need to use $\linCons{}$ like the previous ones, because it modifies the capability in-place, i.e.\ the source register is also the target register.
-The \texttt{jmp} instruction is a simple jump that just sets register $\pcreg$.
-
-%% Sealing
-The operational side of the sealing in \trgcm{} consists of two instructions: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
-% cseal
-Given a sealable $\vsc$ and a set of seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
-% xjmp
-Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them, and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
-A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
-Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
-This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
-Opaque sealed capabilities encapsulate a security domain's local state and authority, and they only become accessible again when control is transferred to the security domain.
-Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one closure with the environment of another.
-% xjmp to unseal
-\trgcm{} does not have an instruction for unsealing capabilities directly, but it can be (partially) simulated using \texttt{xjmp}.
-
-% split/splice
-Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
-For normal capabilities, reduction of authority can be done without actually giving up any authority by duplicating the capability first.
-With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
-In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
-The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities, with $[\var{base},n]$ and $[n+1,\var{end}]$ as ranges of authority.
-Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
-With \texttt{split}, we can reduce the range of authority of a linear capability without losing any authority as we retain it in the second capability.
-The \texttt{splice} instruction essentially does the inverse of \texttt{split}.
-Given two capabilities with adjacent ranges of authority and the same permissions and linearity, \texttt{splice} splices them together into one capability.
-The two instructions work in the same way for seal sets.
-We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
-This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.
-
-The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
-To clarify this, we introduce notions of components and programs from which we construct executable configurations.
-A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
-It has exports that can satisfy the imports of other components.
-% Structure
-A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses (addresses governed by a linear capability).
-The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
-The exports are words each associated with a symbol.
-%base = library
-A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
-The latter can be seen as a program that still needs to be linked with libraries.
-% Linking .
-Components are combined into new components by linking them together, as long as only one is an incomplete program with a main.
-Two components can be linked when their memories, seals, and linear addresses are disjoint.
-They are combined by taking the union of each of their constituents.
-For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
-The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
-
-% Programs
-We can now define the notion of a program as well as a context.
-\begin{definition}[Programs and Contexts]
-  \label{def:program-and-context}
-  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
-  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
-\end{definition}
-% Getting an executable from a component - postpone this to later, but mention:
-% write xor exec
-How a program is initialised to create an executable configuration, is discussed in Section~\ref{sec:form-secur-with}.
-% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
-%Further, we will only consider a specific class of well-formed components, but we will get back to this.
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for details.
-\begin{figure}[tb]
-  \centering
-  \[
-    \begin{array}{rl c rl c rl}
-    s &\in \Symbol & \phantom{make} &
-    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
-    \mathrm{export} &\mathrel{::=} s \mapsto w\\
-    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
-    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
-  \end{array}
-\]
-\vspace{0.5cm}
-\begin{mathpar}
-  \inference{
-    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
-    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
-    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
-    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
-    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
-    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
-    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
-    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
-    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
-    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
-    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
-  } {
-    \comp_0'' = \comp_0 \bowtie \comp_0'
-  }
-  \and
-  \inference{
-    \comp_0'' = \comp_0 \bowtie \comp_0'
-  }{
-    (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \comp_0 \bowtie (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
-  }
-  \end{mathpar}
-% \begin{definition}[Plugging a program into a context]
-%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
-%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
-% \end{definition}
-  \caption{Components and linking of components.}
-  \label{fig:target-component-and-linking}
-\end{figure}
-
-%\FloatBarrier
-% \begin{itemize}
-% \item present our \emph{target language} and its operational semantics (excerpts)
-% \item mention roughly what components look like
-% \end{itemize}
-\section{Linear Stack and Return Capabilities}
-\label{sec:stktokens-explained}
-In this section, we introduce our calling convention \stktokens{} that ensures LSE and WBCF.
-We will gradually explain each of the security measures \stktokens{} takes and motivate them with the attacks they prevent.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-
-      % Stack pointer 1
-      \begin{scope}
-        \clip (4.6,4) rectangle (9,13);
-        \capbracebot{(4.6,4)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 1}
-      \end{scope}
-      Stack pointer 2
-      \begin{scope}
-        \clip (4.8,4) rectangle (9,13);
-        \capbrace{(5.2,9)}{(5.2,13.5)}{adv. stack\\\footnotesize cap. 2}
-      \end{scope}
-
-      \draw[red,thick,->] (4.5,10) to[out=0,in=50] node[midway,right] {} (7.5,7.3);
-      \draw[red,thick,->] (7,7.3) to[out=115,in=0] node[midway,right] {} (4.5,8);
-    \end{tikzpicture}
-    \caption{An adversary uses a previous stack frame's stack pointer.}
-    \label{fig:stack-ptr-abuse}
-  \end{subfigure}
-  \begin{subfigure}{0.18\linewidth}
-    \phantom{testtestes}
-  \end{subfigure}
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-      \fill[red,draw=red,opacity=.4] (-.1,3.9) rectangle (4.6,9.1);
-      \draw[red,thick,->] (4.5,10.5) to[out=0,in=0] node[midway,right,align=center] {return\\ to wrong\\ return ptr} (4.5,3);
-    \end{tikzpicture}
-    \caption{An adversary jumps to a previous stack frame's stack pointer.}
-    \label{fig:ret-ptr-abuse}
-  \end{subfigure}
-
-  \caption{Possible ways to abuse stack and return capabilities.}
-  \label{fig:stack-ret-ptr-abuse}
-\end{figure}
-
-\stktokens{} is based on a traditional single stack, shared between all components.
-To explain the technique, let us first consider how we might already add extra protection to stack and return pointers on a capability machine.
-First, we replace stack pointers with stack capabilities.
-When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
-Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
-% KJAA: Again with the sealed pair. Again, this idea is unclear.
-They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
-
-%informally explain how an adversary may try to abuse stack and return caps
-While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
-Untrusted callees receive a stack capability and a return pair that they are supposed to use for the call.
-However, a malicious callee (which we will refer to as an adversary\footnote{See Section~\ref{sec:well-form-reas} for more details on our attacker model.}) can store the provided capabilities on the heap in order to use them later.
-Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
-In both examples our component and an adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
-The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
-However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
-Generally speaking, we have no reason to ever trust a stack capability received from an untrusted component as that stack capability may have been duplicated and stored for later use.
-In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
-The adversary stores the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed.
-The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.
-
-% Informally explain how we prevent this using linear capabilities
-As the examples illustrate, this naive use of standard memory and object capabilities does not provide sufficient guarantees to enforce LSE and WBCF.
-The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
-Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver no longer has it or a copy of it as it is non-duplicable.} that can be used to prevent this from happening.
-
-% Stack capability linear
-The linear capabilities are put to use by requiring the stack capability to be linear.
-On call, the caller splits the stack capability in two: one capability for their local stack frame and another one for the unused part of the stack.
-The local stack frame capability is sealed and used as the data part of the sealed return pair.
-The capability for the remainder of the stack is given to the callee.
-% Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
-Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
-This means that an adversary would need the stack capability produced by the caller in order to access their local data.
-The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
-This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
-
-% Prevention of attack 2
-In a traditional calling convention with a single stack, the stack serves as a call stack keeping track of the order calls where made in and thus in which order they should be returned to.
-A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
-However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
-This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
-In the presence of adversarial code, we need some mechanism to enforce that the order of the call stack is kept.
-One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
-The callee would have to provide this token on return to prove that it is allowed to return to the caller, and on return the token would be taken back by the caller to prevent it from being spent multiple times.
-As it turns out, the stack capability for the unused part of the stack can be used as such a token in the following way:
-On return the callee has to give back the stack capability they were given on invocation.
-When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack or not.
-They do this by attempting to restore the stack capability from before the call by splicing the return token with the stack capability for the local stack frame which at this point has been unsealed again.
-If the splice is successful, then the caller knows that the two capabilities are adjacent.
-On the other hand, if the splice fails, then they are alerted to the fact that their stack frame may not be the topmost.
-\stktokens{} uses this approach; and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}, it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
-
-      % Stack pointer 1
-      \sealsymb{(4.9,5.5)}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{adv. stack\\\footnotesize cap. 1}
-      % return cap
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{data return\\\footnotesize cap. 2}
-     %  Stack pointer 2
-      \begin{scope}
-        \clip (4.8,4) rectangle (9,13);
-        \lincapbrace{(4.6,9)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 2}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{The non-duplicable linear stack capability for the trusted code's
-      stack frame and the opacity of sealed capabilities ensures LSE.}
-    \label{fig:stack-ptr-abuse-prev}
-  \end{subfigure}
-  \begin{subfigure}{0.18\linewidth}
-    \phantom{testtestes}
-  \end{subfigure}
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-
-      \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
-
-     %  Stack pointer 2
-      \begin{scope}
-        \clip (4.8,4) rectangle (9,13);
-        \lincapbrace{(4.6,9)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 2}
-      \end{scope}
-
-      \redcross{(5,4)}
-    \end{tikzpicture}
-    \caption{The trusted caller fails to splice the stack capability returned by
-    the adversary with the capability for the trusted caller's local stack frame.}
-    \label{fig:ret-ptr-abuse-prev}
-  \end{subfigure}
-  \caption{Abuse of stack and return capabilities prevention.}
-\end{figure}
-
-% Non-empty trusted stack frames.
-In order for a call to have a presence on the call stack, its stack frame must be non-empty.
-We cannot allow empty stack frames on the call stack, because then it would be impossible to tell whether the topmost non-empty stack frame has an empty stack frame on top of it.
-Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in \stktokens{} these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
-This means that a caller using \stktokens{} needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
-There is also a more practical reason for a \stktokens{} caller to make sure their stack frame is non-empty: They need a bit of the stack capability in order to perform the splice that verifies the validity of the return token.
-
-% known stack base
-% + more
-At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
-However, this still does not ensure that the caller's stack frame is on top of the call stack.
-The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
-An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
-This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
-The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
-The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
-At this point (Figure~\ref{fig:stack-base-abuse-b}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
-
-For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In \stktokens{}, the base address of the stack capability is fixed as a compile-time constant (Note: the stack grows downwards, so the base address of the stack capability is the top-most address of the stack).
-The caller verifies the validity of the return token by checking whether the base address of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
-In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
-
-% Where do we get the stack pointer from, how can we know it is linear
-% The other attack.
-% At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
-% However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
-% In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
-% To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
-% The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
-% The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
-% Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
-% At this point, the adversary has everything they need to return from either of the two callbacks.
-% Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
-% From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
-% They are both correct as the adversarial component has created two call stacks, one for each of the components.
-% In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
-% We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
-% A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
-% This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
-
-In \stktokens{}, the stack memory is only referenced by a single linear stack capability at the start of execution.
-Because of this, the return token can be verified simply by checking its base address and splicing it with the caller's stack frame.
-There is no need to check linearity because only linear capabilities to this memory exist.
-
-% Return pointers/return seals
-The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
-Both of the capabilities in the pair are sealed with the same seal.
-% One seal per return point
-All call points need to be associated with a unique seal (a return seal) that is only used for the return capabilities for that particular call point.
-The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that gave rise to the stack frame, breaking WBCF.
-For similar reasons, we cannot allow return seals to be used to seal closures.
-% Don't be stupid
-Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
-This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a capability for reading, either directly or indirectly, a return seal from memory).
-
-% Note about them vs us, we do nothing that they couldn't do
-We have sometimes phrased the description of the \stktokens{} calling scheme in terms of ``them vs us''.
-This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
-However, \stktokens{} is a modular calling scheme: no restriction is put on adversarial components that we do not expect trusted components to meet.
-Specifically, we are going to assume that both trusted and adversarial components are initially syntactically well-formed (described in more detail in Section~\ref{sec:well-form-reas}) which basically just restrict adversarial components to not break machine guarantees initially (e.g.\ no aliases for linear capabilities or access to seals of other components).
-This means that any component can ensure WBCF and LSE by employing \stktokens{}.
-
-% KJAA: ^-- Assuming that there is a global stack address capability, no?
-% Can I "suddenly" begin to impose \stktokens{}?
-
-% Summary of CC
-To summarise, \stktokens{} consists of the following measures:
-\begin{enumerate}
-\item \label{item:check-stkb} Check the base address of the stack capability before and after calls.
-\item \label{item:non-empty-sf} Make sure that local stack frames are non-empty.
-\item \label{item:return-data} Create token and data return capability on call: split the stack capability in two to get a stack capability for your local stack frame and a stack capability for the unused part of the stack.
-The former is sealed and used for the data part of the return pair.
-\item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, every return seal is only used for one call site, and they are not leaked.
-\end{enumerate}
-Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} , except for checking stack base before calls.
-We do not include this check because it only needs to happen once between two calls, so that the check after a call suffices if the stack base is not changed subsequently.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.23\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \actadv{(0,4)}{(4.5,7)}{}
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-      \lincapbracebot{(4.6,4)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-a}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
-      \node[opacity=0.5,rotate=-45] at (2.25,10) {\footnotesize Sent to trusted};
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{}
-
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-      \lincapbracebot{(4.6,7)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-b}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \actadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
-      \node[opacity=0.5,rotate=-45] at (2.25,11) {\footnotesize Returned to adv.};
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{}
-
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{}
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-c}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \actsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (4.6,13.1);
-        \draw[fill=gray!50] (0,9) rectangle (4.5,13.5);
-      \end{scope}
-
-      \lincapbrace{(4.6,2)}{(4.6,7)}{}
-
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{}
-
-      \fill[red,draw=red,opacity=.4] (-.1,6.9) rectangle (4.6,9.1);
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-d}
-  \end{subfigure}
-  \caption{Partial return token used to return out of order.}
-  \label{fig:stk-base-abuse}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-{\footnotesize
-\[
-  \begin{array}{r >{$}p{5.1cm}<{$} r l}
-\multicolumn{2}{l}{\text{// Ensure non-empty stack.}}            & \multicolumn{2}{l}{\text{// Clear tmp registers and jump.}}\\
-1 :& \tmove{\rtmp{1}}{42}                                        &14: & \tmove{\rtmp{1}}{0}\\
-2 :& \tstore{\rstk}{\rtmp{1}}                                    &15: & \txjmp{r_1}{r_2}\\
-3 :& \tcca{\rstk}{(-1)}                                          &\multicolumn{2}{l}{\text{// The following is the return code.}}\\
-\multicolumn{2}{l}{\text{// Split stack in local stack frame and unused.}}      &\multicolumn{2}{l}{\text{// Check that returned stack pointer has base $\stkb$.}}\\
-4 :& \tgeta{\rtmp{1}}{\rstk}                                     &16: & \tgetb{\rtmp{1}}{\rstk}\\
-5 :& \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}                     &17: & \tminus{\rtmp{1}}{\rtmp{1}}{\stkb}\\
-\multicolumn{2}{l}{\text{// Load the call seal.}}                &18: & \tmove{\rtmp{2}}{\pcreg}\\
-6 :& \tmove{\rtmp{1}}{\pcreg}                                    &19: & \tcca{\rtmp{2}}{5}\\
-7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
-8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \rtmp{2}{1}\\
-9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
-\multicolumn{2}{l}{\text{// Seal the local stack frame.}}        &23: & \tfail \\
-10:& \tcseal{\rretd}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Splice with capability for local stack frame.}}\\
-\multicolumn{2}{l}{\text{// Construct code return pointer.}}     &24: & \tsplice{\rstk}{\rstk}{\rdata} \\
-11:& \tmove{\rretc}{\pcreg}                                      &\multicolumn{2}{l}{\text{// Pop 42 from the stack}}\\  
-12:& \tcca{\rretc}{5}                                            &25: & \tcca{\rstk}{1}\\
-13:& \tcseal{\rretc}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Clear tmp register}}\\
-      &    &26: & \tmove{\rtmp{2}}{0}\\
-  \end{array}
-\]
-}
-\caption{
-  % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
-  % Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
-  % Instructions 1-3: ensure that the stack is non-empty.
-  % Instructions 4-5: split stack pointer in two (stack pointer for callee done).
-  % Instructions 6-9: Load the seal designated for this call.
-  % Instruction 10: Seal capability for private stack to create data return pointer.
-  % Instructions 11-13: Seal capability for return point to create code return pointer.
-  % Instructions 14-15: Clean up temporary register and jump.
-  % Instructions 16-26: The return code.
-  % Instructions 16-23: check the stack base.
-  % Instruction 24: join the stack pointer.
-  % Instructions 25-26: pop 42 from the stack.
-  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set of seals it uses and $\offsigma$ the offset in the set of seals to the call seal.
-  $\stkb$ is the globally agreed on stack base.
-  There are some magic numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
-  Line 7: $-5$, offset from line 6 (where $\pcreg$ was copied into $\rtmp{1}$) to line 1.
-  Line 12: $5$, offset to the return address.
-  Line 19: $5$, offset to fail.
-  Line 21: offset to address after fail.
-}
-  \label{fig:call-code}
-\end{figure}
-
-% \begin{itemize}
-% \item informally explain how an adversary may try to abuse stack and return caps
-% \item informally explain how we prevent this using linear capabilities
-% \item use the tikz pictures from the PriSC presentation to explain all of this
-% \end{itemize}
-
-\section{Formulating Security with a Fully Abstract Overlay Semantics}
-\label{sec:form-secur-with}
-As mentioned, the \stktokens{} calling convention guarantees well-bracketed control flow (WBCF) and local state encapsulation (LSE).
-However, before we can prove these properties, we need to know how to even formulate them.
-Although the properties are intuitively clear and sound precise, formalizing them is actually far from obvious.
-
-Ideally, we would like to define the properties in a way that is
-\begin{enumerate}
-\item {\itshape intuitive} \label{def-prop:intuitive}
-\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
-\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
-\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
-\label{def-prop:complete}
-\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
-\end{enumerate}
-
-Previous formalisations in the literature are formulated in terms of a static control flow graph~\cite[e.g., ][]{Abadi2005Theory}.
-While these are intuitively appealing (\ref{def-prop:intuitive}), it is not clear how they can be used to reason about programs (\ref{def-prop:useful}) or other compiler passes (\ref{def-prop:reusable}), they lack temporal safety guarantees (\ref{def-prop:complete}) and do not scale (\ref{def-prop:scalable}) to settings with dynamic code generation (where a static control flow graph cannot be defined).
-\citet{skorstengaard_reasoning_2017} provide a logical relation that
-can be used to reason about programs using their calling convention
-(\ref{def-prop:useful},\ref{def-prop:reusable}), but it is not intuitive (\ref{def-prop:intuitive}), there is no argument for completeness (\ref{def-prop:complete}), and it is unclear whether it will scale to more complex features (\ref{def-prop:scalable}).
-
-We contribute a new way to formalise the properties using a novel approach we call fully abstract overlay semantics.
-The idea is to define a second operational semantics for programs in our target language.
-This second semantics uses a different abstract machine and different run-time values, but it executes in lock-step with the original semantics and there is a very close correspondence between the state of both machines.
-
-The main difference between the two semantics, is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
-Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
-For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.
-
-The fact that \stktokens{} enforces LSE and WBCF is then formulated as a theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
-The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
-Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than components which the trusted component interacts with in the well-behaved overlay semantics.
-In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
-More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics iff they are indistinguishable to arbitrary other components in the overlay semantics.
-
-Our formal results are complicated by the fact that they only hold on a sane initial configuration of the system and for well-behaved components that respect the basic rules of the calling convention.
-For example, the system should be set up such that seals used by components for constructing return pointers are not shared with other components.
-We envision distributing seals as a job for the linker, so this means our results depend on the linker to do this properly.
-As another example, a seal used to construct a return pointer can be reused but only to construct return pointers for the same return point.
-Different seals must be used for different return points.
-Such seals should also never be passed to other components.
-These requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
-The general pattern is that $\stktokens{}$ only protects components that do not shoot themselves in the foot by violating a few basic rules.
-In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components to do certain unsafe things.
-Well-formedness is a requirement for all components (trusted and untrusted), but the reasonability requirement only applies to trusted components, i.e.\ those components for which we provide LSE and WBCF guarantees.
-
-\begin{figure}[b]
-  \centering
-  \[
-    \arraycolsep=1.4pt
-    \begin{array}{rcl}
-      \src{\SealableCaps} & \defbnf& \SealableCaps \mid \src{\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}} \mid\\
-                          & &   \src{\retptrd(\basebnf,\aendbnf)} \mid \src{\retptrc(\basebnf,\aendbnf,\addrbnf)}\\
-      \multicolumn{3}{c}{
-      \begin{array}{lcrclcr}
-        \src{\StkFrame} & \defeq & \src{\Addr \times \MemSeg} & \phantom{skipskipsip} & \src{\Stack} & \defeq & \src{ \StkFrame^*}
-      \end{array}
-                                                                                                                }\\
-      \src{\ExecConf} & \defeq & \Mem \times \Reg \; \src{\times \; \Stack \times \MemSeg} \\
-    \end{array}
-  \] 
-\[
-  \begin{array}{rclcrcl}
-    \src{\Instr} & \defbnf &  \Instr \mid \scall{\offpc,\offsigma}{r}{r}&\phantom{skipskipskip}&
-    \offpc,\offsigma & \in & \nats
-  \end{array}
-\]
-\caption{The syntax of \srccm{}.
-  \srccm{} extends \trgcm{} by adding stack pointers, return pointers, and a built-in stack.
-  Everything specific to the overlay semantics is written in blue.
-}
-  \label{fig:source-syntax}
-\end{figure}
-
-% see e-mail discussion: we do not see a good way to implement this TODO.
-% \todo[inline]{Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.}
-
-\subsection{Overlay Semantics}
-% Describe the things added to the syntax
-The overlay semantics \srccm{} for \trgcm{} views part of the memory as a built-in stack (Figure~\ref{fig:source-syntax}).
-% Built-in stack
-To this end, it adds a call stack and a free stack memory to the executable configurations of \trgcm{}.
-The call stack is a list with all the stack frames that are currently inaccessible because they belong to previous calls.
-Every stack frame contains encapsulated stack memory as well as the program point that execution is supposed to return to.
-The free stack memory is the active part of the stack that has not been claimed by a call and thus can be used at this point of time.
-% Stack pointer
-In order to distinguish capabilities for the stack from the capabilities for the rest of the memory, \srccm{} adds stack pointers.
-A stack pointer has a permission, range of authority, and current address, just like capabilities on \trgcm{}, but they are always linear.
-% Return pointers
-The final syntactic constructs added by \srccm{} are the code and data return pointers.
-The data return pointer corresponds to some stack pointer (which in turn corresponds to a linear capability), and the code return pointer corresponds to some capability with read-execute permission.
-Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to on the underlying machine.
-On \srccm{}, return pointers are generated by calls from the capabilities they correspond to on \trgcm{}, and they are turned back to the capabilities they correspond to upon return.
-% Even though the return pointers both correspond to capabilities with read permission, neither can be used for reading. 
-% Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack memory that is encapsulated in a stack frame which would break LSE.
-
-The opaque nature of the return pointers is reflected in the interpretation of the instructions common to both \trgcm{} and \srccm{} as \srccm{} does not add special interpretation for them in non-\texttt{xjmp} instructions.
-Stack pointers, on the other hand, need to behave just like capabilities, so \srccm{} adds new cases for them in the semantics, e.g.\ \texttt{cca} can now also change the current address of a stack pointer as displayed in Figure~\ref{fig:source-op-sem}.
-Similarly, \texttt{load} and \texttt{store} work on the free part of the stack when provided with a stack pointer.
-A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics has on the underlying machine.
-In other words, there should only be stack pointers for the stack memory.
-
-% Back to high-level view
-As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
-Untrusted components are not assumed to do so.
-To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
-Only instructions at these addresses can be interpreted as the \srccm{} native call and push frames to the call stack which guarantees LSE and WBCF.
-The constant $\ta$ is a parameter of the \srccm{} step relation.
-Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
-
-Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
-This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
-The sequence of instructions have to correspond to a call, i.e.\ the
-instructions in Figure~\ref{fig:call-code} ({\footnotesize  $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is always $26$, i.e.\ the number of instructions).
-Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.
-
-% Call/return
-The interpretation of $\scall{\offpc,\offsigma}{r_1}{r_2}$ is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following:
-The registers $r_1$ and $r_2$ are expected to contain a code-data pair sealed with the same seal and the unsealed values are invoked by placing them in the $\pcreg$ and $\rdata$ registers, respectively.
-The current active stack and the stack capability are split into the local stack frame of the caller and the rest.
-$\scall{}{}{}$ also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
-The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of sealed return capabilities.
-The return capabilities are sealed with the seal designated for the call.
-
-The return capabilities, $\retptrc$ and $\retptrd$ are sealed and can only be used using the \texttt{xjmp} instruction, to perform a return.
-When this happens, the topmost call stack frame $(\opc,\ms_\var{local})$ is popped from the call stack.
-In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must match the domain of the local stack.
-If the return succeeds, the stack pointer is reconstructed, and the local stack becomes part of the active stack again.
-
-\srccm{} supports tail calls.
-A tail call is a call from a caller that is done executing, and thus doesn't need to be returned to or preserve local state.
-This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in call.
-To perform a tail call, the caller simply transfers control to the callee using \texttt{xjmp}.
-The tail-callee should return to the caller's caller, so the caller leaves the return pair they received for the callee to use.
-
-It is important to observe that the operational semantics of \srccm{} natively guarantee WBCF (well-bracketed control flow) and (local stack encapsulation) for calls made by trusted components.
-By inspecting the operational semantics of \srccm{}, we can see that it never allow reads or writes to inactive stack frames on the call stack.
-The built-in call for trusted code pushes the local stack frame to the inactive part of the stack, together with the return address. 
-Such frames can be reactivated by \texttt{xjmp}ing to a return capability pair, but only for the topmost stack frame and if the return address corresponds to the one stored in the call stack.
-In other words, WBCF and LSE are natively enforced in this semantics.
-
-\begin{figure}[htb]
-  \centering
-  \begin{mathpar}
-    \inferrule{\src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
-               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
-               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
-               \src{\perm \in \{\rwx,\rx\}} \\
-               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2} \cdots \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
- }
-              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} }
-  \end{mathpar}
-  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.6cm}|}
-    \hline
-    i \in \src{\Instr}                                 & $\sem{i}(\Phi)$ & Conditions\\
-    \hline
-    \halt                                        & $\halted$ & \\
-    \hline
-    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
-    \hline
-    \store{r_1}{r_2}                             & $\srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray$  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
-    \hline
-    \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
-    \hline
-    \arraycolsep=0pt\array{c}
-    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}}
-    \endarray &
-        $\srcalt{\var{xjmpRes}(c_1,c_2,}$
-        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1,r_2}{w_1,w_2}\\
-            &\updReg{\rretc}{s_c}\\
-            &\updReg{\rretd}{s_d}\\
-            &\updReg{\rstk}{c_\stk}\\
-            &\update{\ms_\stk}{\ms_{\stk,\var{rest}}}\\
-            &\update{\stk}{\stk'}\\
-            \endarray\right))}$
-       &
-      \srcalt{
-         $\begin{multlined}
-           \ms_{\stk,\var{local}},c_\var{local},\ms_{\stk,\var{rest}},c_\stk =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
-         \end{multlined}$
-    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
-    $\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
-    $\sigma =\begin{multlined}[t]
-      \var{getCallSeal}(\\
-      \Phi.\reg(\pcreg), \Phi.\mem,\offpc,\offsigma)  \text{ and }
-    \end{multlined}$
-    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\opc,c_\var{local})$ and
-    $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
-                        $\Phi.\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
-                        }\\
-    \hline 
-    \multicolumn{3}{|c|}{\dots} \\
-    \hline
-    \_                                           & $\failed$ & \totherwise \\
-    \hline
-  \end{tabular}
-\begin{multline*}
-  \srcxjmpres{c_1,c_2,\Phi} = \\
-  \left\{ 
-    \arraycolsep=1.4pt
-    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
-      \arraycolsep=0pt
-    \begin{array}{rl}
-      \Phi & \updReg{\pcreg}{c_1}\\
-           & \updReg{\rdata}{c_2}
-    \end{array}
-    &\phantom{mak} & $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
-      \arraycolsep=0pt
-      \array[c]{rl}
-      \Phi&\updReg{\pcreg}{c_\opc}\\
-          &\updReg{\rstk}{c_\stk}\\
-          &\updReg{\rdata}{0}\\
-          &\update{\stk}{\stk'}\\
-          &\update{\ms_\stk}{\ms_\stk \uplus \ms_\var{local}}
-      \endarray
-           & &
-               $\arraycolsep=0pt\array{l}
-        \src{(\opc, \ms_\var{local}) :: \stk'= \Phi.\stk \tand} \\
-      \src{c_1 = \retptrc(\baddr,\eaddr,\opc)}\\
-      \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge \dom(\ms_\var{local}) = [\aaddr_\stk,\eaddr_\stk]}\\
-      \src{c_\stk = \var{reconstructStackPointer}(\Phi.\reg(\rstk),c_2) \tand}\\
-      \src{c_\opc =  ((\rx,\normal),\baddr,\eaddr,\opc)} 
-      \endarray$\\
-      \failed &  & \totherwise
-    \end{array} 
-\right.
-\end{multline*}
-  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted). Auxiliary definitions are found in Figure~\ref{fig:source-op-sem-aux}. }
-  \label{fig:source-op-sem}
-\end{figure}
-
-\begin{figure}
-    \begin{multline*}
-      \src{\var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk) = \ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk} \mathit{\ iff\ } \\
-      \left\{ 
-      \begin{array}{l}
-      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
-      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
-      \src{\ms_{\stk,\var{unused}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]} }\\
-      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
-      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
-      \end{array}
-      \right.
-    \end{multline*}
-    \begin{align*}
-      \src{ \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} &\ \src{= \opc, c_\opc} \mathit{\ iff\ }
-      \left\{ 
-      \begin{array}{l}
-      \src{\opc = \aaddr + \calllen \wedge{}}\\
-      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
-      \end{array}
-       \right.\\
-      \src{\var{getCallSeal}(c_\pcreg,\mem,\offpc,\offsigma)} &\ \src{= \sigma} \mathit{\ iff\ }
-      \left\{
-      \begin{array}{l} 
-      \src{c_\pcreg = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
-      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge{} \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \wedge{}}\\
-      \src{\sigma = \sigma_\aaddr + \offsigma }
-      \end{array}
-      \right.\\
-     \src{\var{sealReturnPair}(\sigma,c_\opc,c_\var{local})} &\ \src{= \sealed{\sigma,c_\opc},\sealed{\sigma,c_\var{local}}}
-    \end{align*}%
-  % \begin{multline*}
-  %   \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}\mathit{\ iff}\\
-  %     \quad\quad\left\{ 
-  %       \begin{array}{l}
-  %         \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)}\\
-  %     \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{rest}},c_\stk,c_\var{local\_data} = \var{splitStack}(\reg, \ms_\stk) \wedge{}}\\
-  %     \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
-  %     \src{\stk' = (\opc,\ms_{\stk,\var{local}}) :: \stk \wedge{}}\\
-  %     \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
-  %     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local\_data},c_\opc) \wedge{}}\\
-  %     \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
-  %     \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
-  %     \end{array}\right.
-  % \end{multline*}
-  \begin{multline*}
-    \src{\var{reconstructStackPointer}(\stkptr{\rw, \stkb, \aaddr_\stk-1,\_},\retptrd(\aaddr_\stk,\eaddr_\stk)) =}\\ \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk}} \mathit{\ iff\ }
-        \src{\stkb \leq \addr_\stk }
-   \end{multline*}
-
-  \caption{Auxiliary definitions used in the operational semantics of \srccm{}.}
-  \label{fig:source-op-sem-aux}
-\end{figure}
-% \begin{figure}
-% % \[
-% %     \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
-% % \]
-% % \[
-% %     \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
-% %     \mathit{\ iff\ }
-% %       \left\{ \array{l}
-% %         \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
-% %         \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
-% %         \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
-% %         \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
-% %         \endarray
-% %      \right.
-% % \]
-% % \[
-% %   \src{\var{retPtrForMatchLocalStack(\retptrd(\aaddr_\stk,\eaddr_\stk),\ms_\var{local})}} \mathit{\ iff\ } \src{\dom(\ms_\var{local} \subseteq [\aaddr_\stk,\eaddr_\stk])}
-% % \]
-
-% % \begin{multline*}
-% %     \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
-% %       \quad\quad\left\{ 
-% %       \begin{array}{l}
-% %       \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
-% %       \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
-% %       \src{c_\opc =  \var{tokenToPc}(c_1)} \\
-% %       \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
-% %       \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
-% %       \end{array}\right.
-% %   \end{multline*}
-
-%   \caption{The return condition in the operational semantics of \srccm{}.}
-%   \label{fig:source-op-sem-return}
-% \end{figure}
-\subsection{Well-Formedness and Reasonability}
-\label{sec:well-form-reas}
-% Well-formed components
-% \begin{figure}[htb]
-%   \centering
-%   \begin{mathpar}
-%   \inference{
-%     \dom(\mscode) = [\baddr,\eaddr]&
-%     [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
-%     \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
-%     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
-%     \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
-%     A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-%     \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
-%     \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} & \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}}& \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
-%     \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}\\
-%     \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
-%     \dom(\msdata) \mathrel{\#} \ta
-%   }{
-%     \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
-%   }
-%   \and
-%   \inference{
-%     \comp_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
-%     \vdash \comp_0 & (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
-%   }{
-%     \vdash (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
-%   }
-% \end{mathpar}
-%   \caption{Exceprt of well-formedness judgement}
-%   \label{fig:well-formedness}
-% \end{figure}
-% Well-formedness highlights:
-% Main exported
-% Code memory:
-% - Only instructions and seals
-% - calls associated with return seals unique return seal
-% - only uses seals specified for components
-% - (No linear access)
-% Data memory:
-% - Data
-% - Capabilities fora data memory (linearity respected)
-% - Sealed caps sealed with clos seal
-% Exports
-% - Any Word valid in data memory or
-% - Sealed capability for code (sealed with clos seal)
-% Imports
-% - Address must be in data memory
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
-We elide the details of the judgement here and describe the main points:
-For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
-As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
-The return seals $\sigrets$ are the seals supposed to be used to seal return pointers and the closure seals $\sigcloss$ all the other seals in a component.
-If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjoint from the set of trusted addresses $\ta$, then there should be no return seals.
-The code memory may contain sets of seals but only for the seals in $\sigrets$ and $\sigcloss$.
-Other than that, code memory only contains instructions in the form of integers.
-When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), the call must have access to the return seal it specifies.
-The return seal also needs to be unique to that call; that is, no other call can specify the same seal as its return seal.
-The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
-The capabilities in the data memory can only have authority over the data memory itself.
-This allows components to have initial data structures.
-In order to respect Write-XOR-Execute, the capabilities in data memory cannot have execute permission.
-The capabilities also need to respect linearity which means that linear capabilities cannot be aliased by any other capability, and they must be for a range of the predetermined linear addresses $A_\linear$.
-The sealed capabilities must be sealed with a closure seal, and the sealable can be anything that is allowed to reside in data memory.
-The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
-
-% Reasonability
-The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
-In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
-More precisely, we expect four things of a reasonable component: (1) It checks the stack base address before performing a call.
-As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
-(2) It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
-(3) It does not leak return and closure seals or means to retrieve them.
-This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
-There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
-(4) It does not store return and closure seals or means to get them.
-By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
-We elide the details here and refer to~\citet{technical_report}
-
-In our result, we assume that adversarial components are well-formed, but not necessarily reasonable.
-The well-formedness assumption ensures that the trusted component can rely on basic security guarantees provided by the capability machine.
-For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
-The adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
-Can adversarial code call into trusted components?
-The answer to that question is yes but not with LSE and WBCF guarantees.
-Formally, adversarial code can contain the instructions that constitute a call.
-However, for untrusted code, \srccm{} will not execute those instructions as a "native call" but execute the individual instructions separately.
-The callee then executes in the same stack frame as the caller, so WBCF and LSE do not follow (for that call).
-
-We will assume trusted components, for which WBCF and LSE are guaranteed, to be both well-formed and reasonable. 
-%The reasonability definition has also been left out for space reasons, and we refer to the technical report~\citep{technical_report} for full definitions.
-
-% \begin{definition}[Reasonable configuration]
-%   \label{def:reasonable-conf}
-%   We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
-%   iff for $n' \leq n$:
-%   \begin{itemize}
-%   \item \emph{Guarantee stack base address before call...}
-%   \item \emph{Use return seals only for calls, use closure seals appropriately...}
-%   \item \emph{Don't store private stuff...} If $\src{\Phi}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then
-
-%     Then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ steps.
-%   \item \emph{Don't leak private stuff...}
-%   \end{itemize}
-% \end{definition}
-% Reasonable configuration
-% Reasonable component
-
-\subsection{Full Abstraction}
-All that is left before we state the full-abstraction theorem is to define how components are combined with contexts and executed, so that we can define contextual equivalence.
-
-\begin{figure}
-  \centering
-  \begin{equation*}
-    \inference{
-      c_{\mathrm{main},c},c_{\mathrm{main},d}  = \sealed{\sigma, c_{\mathrm{main},c}', c_{\mathrm{main},d}'} &
-      \nonExec{c_{\mathrm{main},d}'}&
-      \reg(\pcreg,\rdata) = c_{\mathrm{main},c}', c_{\mathrm{main},d}' \\
-      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
-      \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
-      \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0&
-      \range{\ms_\stk} = \{0\}&
-      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
-      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
-      \overline{\var{import}} = \emptyset
-    }{
-      ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
-    }
-  \end{equation*}
-  \caption{The judgement $\var{prog} \rightsquigarrow \Phi$, which defines the initial execution configuration $\Phi$ for executing a program $\var{prog}$.}
-  \label{fig:init-ec}
-\end{figure}
-
-% How a program is lifted to an executable configuration
-Given a program $\var{comp}$, the judgement $\var{comp} \rightsquigarrow \Phi$ in Figure~\ref{fig:init-ec} defines an initial execution configuration that can be executed.
-It works almost the same on \trgcm{} (conditions in \trg{red}) and \srccm{} (conditions in \src{blue}).
-On both machines a stack containing all zeroes is added, as part of the regular memory on \trgcm{} and as the free stack on \srccm{}.
-On \srccm{}, the initial stack is empty as no calls have been made.
-The component needs access to the stack, so a stack pointer is added to the register file in $\rstk$.
-On \trgcm{} this is just a linear read-write capability, but on \srccm{} it is the representation of a stack pointer.
-The entry point of the program is specified by main, so the two capabilities are unsealed (they must have the same seal) and placed in the $\pcreg$ and $\rdata$ registers.
-Other registers are set to zero.
-
-Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
-%The remaining defintions
-\begin{definition}[Plugging a component into a context]
-  \label{def:comp-context-plugging}
-  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
-  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
-\end{definition}
-\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
-  \label{def:contextual-equivalence}
-  ~
-  \begin{description}
-  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
-  \begin{equation*}
-    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
-  \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
-  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
-  \begin{equation*}
-    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
-  \end{equation*}
-\end{description}
-where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
-\end{definition}
-With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
-We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
-This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
-% Full abstraction thm
-% \todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
-\begin{theorem}
-  \label{thm:full-abstraction}
-  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
-  \begin{gather*}
-    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2} \qedhere
-  \end{gather*}
-\end{theorem}
-Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
-Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
-This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
-If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
-In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
-In order to enforce these properties, components on \trgcm{} use \stktokens{}.
-Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
-% \todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}
-
-% \begin{itemize}
-% \item present our source language, its operational semantics (excerpts)
-% \item tell more about components, specifically well-formed according to judgement
-% \item mention our assumption of reasonability
-% \item present the full abstraction theorem.
-% \end{itemize}
-
-\section{Proving full abstraction}
-\label{sec:fa-proof}
 %\todo[inline]{Write small introduction.
@@
 To prove Theorem~\ref{thm:full-abstraction}, we will essentially show that trusted components in \srccm{} are related in a certain way to their embeddings in \trgcm{}, and that untrusted \trgcm{} components are similarly related to their embeddings in \srccm{}.
-We will then prove that these relations imply that the combined programs have the same observable behavior, i.e.\ one terminates iff the other does.
-The hard part is in defining when components are related.
+We will then prove that these relations imply that the combined programs have the same observable behavior, i.e.\ one terminates if and only if the other does.
+The difficult part is to define when components are related.
 In the next section, we give an informal overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
 
-\subsection{Kripke Worlds}
+\subsection{Kripke worlds}
 \label{subsec:worlds}
@@
 However, values are often only related under certain assumptions about the rest of the system.
-For example, the linear data part of a return capability should only be related to the corresponding \srccm{} capability if no other value in the system references the same inactive stack frame and if it is sealed with a seal that is only used for return pointers to the same code location.
-To accomodate such conditional relatedness, we construct our relation as a step-indexed Kripke logical relation with recursive worlds.
-Space constraints prevent us from explaining this in full detail in this paper. 
-Instead, we will only highlight specific important parts of the logical relation in this section, and we refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
-Additionally, for presentation, we will omit details, like step indexing, that are important for correctness but otherwise uninteresting.
-
+For example, the linear data part of a return capability should only be related to the corresponding \srccm{} capability if no other value in the system references the same inactive stack frame and it is sealed with a seal only used for return pointers to the same code location.
+To accommodate such conditional relatedness, we construct the relation as a step-indexed Kripke logical relation with recursive worlds.
+\begin{jversion}
+In this presentation, we try to include most of the details.
+We refer to the conference version of this paper for a presentation\lau{insert reference} with fewer details and emphasis on the intuition.
+\end{jversion}
+
+\begin{jversion}
 Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds.
+A world is a semantic model of the memory.
+In its simplest form, it is a collection of invariants that the memory must satisfy.
+The invariants of a world can vary in complexity and expressiveness depending on the application.
+The possible contents of the memory also influences what the world looks like.
+For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} are modelled by the worlds.
+In order to related \trgcm{} and \srccm{}, we need to model all the features of \srccm{} in the worlds which means we have to model:
+\begin{itemize}
+\item Three kinds of memory: heap, stack of local frames, and free stack
+\item Linearity
+\item Call stack
+\item Seals
+\end{itemize}
+
+\subsubsection{Triple world and regions}
+\srccm{}s memory is split in three: heap, free stack and encapsulated local stack frames.
+In order to model the three kinds of memory, we simply have three sub-worlds.
+That is, our world is defined as
+\[
+  \World = \Worldh \times \Worlds \times \Worldfs
+\]
+The sub-worlds are partial maps from names $\RegName$, modelled as natural numbers, to invariants.
+In order to define what this actually means, we need to be more precise about what we mean by invariants.
+We call the invariants regions because, as we will see later, they turn out to not be invariant.
+A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments ($\Rel{\MemSeg \times \MemSeg}$).
+Intuitively, we want to be able to say that two memory segments are related when their contents are related.
+That is, for every address in the two memory segments, the words that resides there must be related.
+In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now consider the intuition.
+Say we have two integers, then they are related if they are equal.
+If we instead have two capabilities, then they should also be related if they in some sense are equal, but what does it mean for capabilities to be ``equal''?
+Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably do the same computation.
+As a capability points to a piece of memory, the authority of the capability depends on the contents of that memory.
+To say whether two capabilities are related, we must know the possible contents of the memory.
+This is exactly what our world express, so our regions will be world indexed, i.e.
+\[
+  \Worldh = \RegName \parfun (\World \fun \Rel{\MemSeg \times \MemSeg})
+\]
+At this point, we can see that we have constructed a recursive domain equation.
+If we inline $\Worldh$ in $\World$, then we have a circular equation with no solution because the self-reference happens in a negative position.
+Luckily, we can solve circular equations if we move to a different domain.
+For now, we will ignore the problem and return to the issue in Section~\ref{subsubsec:rec-dom-eq}.
+
+For the sake of readability, we introduce the following notation
+\begin{gather*}
+  \pwheap = \pi_1(W)\\
+  \pwpriv = \pi_2(W)\\
+  \pwfree = \pi_3(W)
+\end{gather*}
+
+If we just wanted to model the memory, then the above regions would suffice, but we also need the world to model linearity and seals.
+In the following subsections, we will gradually introduce the necessary components for our worlds.
+
+\subsubsection{Linearity and world joins}
+% Linearity
+The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they can reference.
+In order to model this uniqueness, we need to keep track of which parts of memory that are uniquely referenced and make sure that only one capability references the unique parts.
+%% Bookkeeping done by worlds
+We use the world to keep track of what parts of memory must be uniquely referenced by having two kinds of regions: shared and spatial.
+If a memory segment is governed by a shared region, then non-linear capabilities may reference it.
+On the other hand, if a memory segment is governed by a spatial region, then only linear capability may reference it.
+%%% spatial/spatial_owned
+We cannot let multiple linear capabilities reference the same memory, so we add ownership to spatial regions.
+It is only the spatial owned regions that can be referenced by a capability.
+Specifically, we add tags $\spatial$ and $\spatialo$ to the spatial regions:
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})\uplus \\ 
+  \end{array} \right.
+\]
+For readability, we also add a tag $\pure$ to the shared regions:
+\[
+  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
+\]
+
+% What parts of memory are linear
+We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we have introduced enough to finish the definition of each of the three sub-worlds.
+The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and non-linear capabilities, so it should contain both shared and spatial regions.
+For this reason, it is defined as
+\[
+  \Worldh = \RegName \parfun (\World \fun \Regionh + \Regions)
+\]
+\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
+For instance the $\Worldfs$ is defined as
+\[
+  \Worldfs = \RegName \parfun (\World \fun \Regions)
+\]
+$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
+$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
+Conceptually, this means that each of the stack frames is connected with a return point in some code.
+In a traditional C calling convention, the code return point would even be stored in the stack frame.
+However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
+Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together constitute the sealed return pair.
+In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
+To this end, each shared region in $\Worlds$ must be paired with a return address:
+\[
+\Worlds = \RegName \parfun (\World \fun \Regions) \times \Addr 
+\]
+
+% Ensuring linearity
+The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
+The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities uniquely references part of memory.
+
+\subsubsection{Seals}
+\label{subsubsec:seals}
+% StkTokens designate seals for certain purposes, return, closure, etc
+In order to guarantee well-bracketed control flow and local-state encapsulation, \stktokens{} designates seals for return capabilities and closure seals and relies on designated seals only being used for their intended purpose.
+The designation of return seal is a system invariant that is modelled in the world as a \textit{seal interpretation function}.
+This function takes a seal and returns a world-indexed relation that relates all the sealables that may be sealed with this specific seals.
+That is
+\[
+  \Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps}
+\]
+%% Shared region has a seal indexed invariant that specifies what sealables can be sealed with a specific seal
+In Section~\ref{subsubsec:ft-and-revocation}, we will see that not all regions are invariant.
+The seal interpretation function should be an invariant, so we need to add it to a type of region that is invariant.
+As we will see, shared regions are invariant, so we add the seal interpretation function as follows
+\begin{multline*}
+  \Regionh = 
+  \{\pure \} \times (\World \fun \Rel{\MemSeg\times\MemSeg}) \times \\
+  (\Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps})
+\end{multline*}
+
+\lau{Maybe add a bit more of explanation? Invariant needed for the LR to make sure that designated seals are used for their purpose. }
+
+\subsubsection{Future worlds and revocation}
+\label{subsubsec:ft-and-revocation}
+% Memory evolves over time, we need to model this.
+The world specifies the contents of the memory.
+When a machine executes, the contents of the memory changes and parts of memory may be repurposed.
+The world must model these changes.
+% Future World
+To this end, we have a future world relation $\future$ that describes how worlds may change over time.
+For our purpose, relating \trgcm{} to \srccm{}, we are mainly interested in changes to the stack as it is the part that will be repurposed.
+Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions that govern call frames that have been popped from the call stack, so the stack frame can be reused for a new call frame.
+We allow introduction of new regions simply by extension.
+That is, a future world can have new regions that was not present in a past world.
+Things are a bit more complicated when it comes to revocation.
+\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).}
+For technical reasons, we need the worlds to be extensional which means that when it comes to removing regions in future worlds, we cannot simply forget about them.
+Instead of removing the regions, we keep them around and tag them as removed with a $\revoked$ tag.
+This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.
+It is not all regions that need to be able to be revoked.
+Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities in some sense can be revoked.
+A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory.
+This means that it is safe to repurpose that part of memory.
+On the other hand, non-linear capabilities may have an alias, so it is not safe to repurpose the memory a non-linear capability has authority over.
+It is only spatial regions that specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
+The $\revoked$ region does not specify any memory, so we drop the world indexed relation.
+That is, we have
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\revoked \}
+  \end{array} \right.
+\]
+The rules for the future region relation are displayed in Figure~\ref{fig:ft-reg-rel}.
+% Revocation of spatial regions. 'revocation' of linear capabilities.
+All regions are future regions of themselves, so regions are allowed to stay unchanged.
+The $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
+% Spatial -> spatialo ~ affine ?
+\todo[inline]{Lau: explain why $\spatialo \future \spatial$ gives us affine capabilitites.
+This is claimed in the technical report, but I don't see how this is different from revoking $\spatial$ regions and adding a new region with the future world relation.}
+\begin{figure}[htb]
+  \centering
+  \begin{mathpar}
+  \inferrule{ }{ \revoked \future (\spatial,\_)}
+  \and
+  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
+  \and
+  \inferrule{ r \in \Regions \cup \Regionh }{ r \future r}
+\end{mathpar}
+  \caption{Future region relation.}
+  \label{fig:ft-reg-rel}
+\end{figure}
+
+With the future region relation in place, we define the future world relation as follows:
+\begin{mathpar}
+  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r) }
+            { W' \future W }
+\end{mathpar}
+The relation says that each of the three worlds must be an extension of the past world and for each of the existing regions that must be a future region.
+% injective function
+Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
+The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard:esop18} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.
+
+% TODO monotonicity of regions
+
+\subsubsection{Solving the recursive domain equation}
+\label{subsubsec:rec-dom-eq}
+%solving the recursive domain equation means that we can present the actual definitions
+In the previous sections, we have sketched what we would like our worlds to look like.
+However, the worlds we want constitute a self-referential domain equation for which no solution exists in set theory.
+% Move to a different domain
+Therefore, we need to move to a different domain with enough structure for such an equation to have a solution.
+Specifically, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
+Intuitively, a c.o.f.e. is a set with some extra structure.
+Specifically, c.o.f.e.'s have a step-index which provide sufficient structure for us to solve recursive domain equations.
+In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.
+
+% Cofe's
+In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
+We loosely follow the presentation in \citet{Birkedal:tutorial-notes}.
+\begin{definition}[Ordered family of equivalences (o.f.e.)]
+An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and for each $n$ $\nequal$ is an equivalence relation.
+The pair must satisfy the following properties
+\begin{enumerate}
+\item $\nequal[0]$ is the total relation.
+\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
+\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
+\end{enumerate}
+\end{definition}
+\noindent One can think of the indexes on the family of equivalences as a measure of precision.
+The larger $n$ is, the more refined the equivalence may be.
+On the other hand as $n$ decreases, the equivalence becomes more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
+At index 0, all precision is lost and the equivalence cannot distinguish anything.
+% TODO: Lau write intuition for sequences and why they are necessary.
+\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
+  \label{def:cauchy-sequence}
+  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
+  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
+  in $X$.
+
+  Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
+  elements of $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
+  \begin{align*}
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
+  \end{align*}
+  or in words, the elements of the chain get arbitrarily close.
+
+  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
+  \begin{align*}
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
+  \end{align*}
+\end{definition}
+Functions between c.o.f.e.'s must retain the added structure relatively to sets.
+To this end, we require all our functions to be \emph{non-expansive} which means that $n$-equivalences are preserved.
+A function is called \emph{contractive} when it not only retain equivalences but makes them more precise.
+\begin{definition}
+  \label{def:nonexpansive-contractive-ofe}
+  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
+  We 
+  The function $f$ is
+  \begin{description}
+  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
+\[
+  x \nequal_X x' \implies f(x) \nequal_Y f(x')
+\]
+  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
+\[
+  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
+  \qedhere
+\]
+  \end{description}
+\end{definition}
+\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
+However, this is not the only structure we would like to impose on our sets.
+In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
+The future world relations is essentially a preorder structure that we would like to impose on our sets.
+We impose this structure on our sets by moving to \emph{preordered c.o.f.e.'s}.
+% preordered cofe
+\begin{definition}[Preordered c.o.f.e.]
+  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
+  \begin{itemize}
+  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
+  \end{itemize}
+\end{definition}
+The new structure we add to the set, the preorder, must be consistent with the existing structure which is why it must preserve limits.
+
+Functions between preordered c.o.f.e's must preserve the ordering which means that they must be monotone with respect to the ordering.
+% monotone functions
+\begin{lemma}
+  \label{def:monotone-preordered-cofe}
+  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty},\future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} ,\future_Y\right)$ be two preordered c.o.f.e.'s and $f$ a function from the set $X$ to the set $Y$.
+  We 
+  The function $f$ is
+  \begin{description}
+  \item[monotone] when for all $x, x' \in X$,
+\[
+  x' \future_X x \implies f(x) \future_Y f(x')
+\]
+  \end{description}
+\end{lemma}
+% Not sure this is necessary:
+% % preordered cofe constructions
+% %% cofe to preordered cofe
+% \begin{lemma}[c.o.f.e.'s are preordered c.o.f.e's]
+%   A c.o.f.e $\cofe{X}$, is also a preordered c.o.f.e. with the order $=$.
+% \end{lemma}
+In order to express the domain equation, we must lift all parts of the world definition from sets to c.o.f.e.'s.
+That is, we need to lift the sets of partial functions, monotone non-expansive functions, and relations, respectively to c.o.f.e.'s.
+
+We need to lift the set of relations over $R$, $\Rel{R}$, to a c.o.f.e.
+However, the set of relations have no inherent step-index which means that we need to add one.
+We need to add the step-index in such a way that we can define a family of equivalences that satisfies the c.o.f.e.\ conditions.
+To this end, we define the uniform relations over $R$ $\URel{R}$.
+%% URel
+\begin{definition}[Uniform relation]
+  Given a relation $R \subseteq X \times Y$, we define $\URel{R} \subset \powerset{\nats \times R}$ as
+  \[
+    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats\ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp (n,r) \in A\}
+  \]
+  The c.o.f.e.\ $\cofe{\URel{R}}$. Define erasure as
+  \[
+    \erase{A}{n} \defeq \{(k,a) \in A \mid k < n\}
+  \]
+  and the family of equivalences as
+  \[
+    A \nequal B \text{ iff } \erase{A}{n} = \erase{B}{n}
+  \]
+\end{definition}
+The $n$-equivalences of the uniform relation basically forgets everything from the step and up and requires the rest to be equal.
+This means that everything is equal at step $0$ as everything is left out.
+
+Next, we define the preordered c.o.f.e.\ of monotone and non-expansive functions.
+%% Monotone, non-expansive
+\begin{lemma}[Preordered c.o.f.e. of monotone and non-expansive functions]
+  Given preordered c.o.f.e.\ $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future_W\right)$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ as follows
+  \[
+    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
+  \]
+  and preorder
+  \[
+    h \future h' \text{ iff } \forall r \in \dom(h) \ldotp h(w) = h'(w)
+  \]
+\end{lemma}
+Next, we define the product preordered c.o.f.e.
+%% Product
+\begin{lemma}[Product preordered c.o.f.e.]
+  Two preordered c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ can be composed to a product preordered c.o.f.e $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
+  where the family of equivalences are defined as follows
+  \[
+    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
+  \]
+  and the order as follows
+  \[
+    (x,y) \future (x',y') \text{ iff } x \future_X x' \wedge y \future_Y y'
+  \]
+\end{lemma}
+We use the product preordered c.o.f.e.\ when we just need to compose two existing preordered c.o.f.e.'s into a new preordered c.o.f.e.
+However, in the case of our regions, this do not suffice as they need to use the future region relation we defined in Section~\ref{subsubsec:ft-and-revocation}.
+We therefore define separate preordered c.o.f.e.'s that we use to define these.
+\begin{lemma}[$\Regions$ preordered c.o.f.e.]
+  Given preordered c.o.f.e.\ $\left(H, \left(\nequal_H\right)_{n=0}^{\infty}, \future_H \right)$ define $\left(\{\spatial\} \times H + \{\spatialo\} \times H + \revoked, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
+  where $\future$ is the future region relation from Section~\ref{subsubsec:ft-and-revocation} and
+  \[
+    r \nequal r' \text{ iff } n=0 \vee \left\{
+      \begin{aligned}
+      (\_,h),(\_,h') \in \{ \spatial  \} \times H \wedge h \nequal h' \vee \\
+      (\_,h),(\_,h') \in \{ \spatialo \} \times H \wedge h \nequal h' \vee \\
+      r,r' = \{revoked\}      
+    \end{aligned}\right.
+  \]
+\end{lemma}
+
+%% Sum
+Sum c.o.f.e.'s are defined in a similar manner.
+
+\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
+%% Partial function from set
+\begin{lemma}[Preordered c.o.f.e.\ of partial functions]
+  Given a set $S$ and a preorder c.o.f.e $P$, the set of partial functions $S \parfun P$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
+  \[
+    f \nequal[0] g \wedge f \nequal g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \nequal g(x)
+  \]
+  and preorder
+  \[
+    f \future g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \future g(x)
+  \]
+\end{lemma}
+
+%% Later cofe
+\begin{lemma}[$\blater$ preordered c.o.f.e.]
+  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
+  \[
+    \blater\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right) = \left(X, \left(\nequal[n]'\right)_{n=0}^{\infty}, \future \right)
+  \]
+  where
+  \[
+    x \nequal[0]' x' \wedge x \nequal[n]' x' \texttt{ iff } x \nequal[n+1] x'
+  \]
+\end{lemma}
+
+% partial functions
+%
+
+% New world definitions
+The world definitions we presented so far needs to move to the domain of preordered c.o.f.e's.
+Morally, the definitions are the same, though.
+The definition of the two kinds of regions now uses $\URel{}$ rather than $\Rel$, and the world indexed-functions are monotone.
+Finally, all functions are now non-expansive.
+\begin{multline*}
+  \Regionh = 
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
+\end{multline*}
+and
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
+    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
+    \{\revoked\}
+  \end{array} \right.
+\]
+
+% Reference to solution?
+%% Aleš note? Tutorial note?
+%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
+
+% Argue, informally, that it is okay
+
+% Solution theorem
+
+\subsubsection{Joining worlds}
+
+In order to express this, we will need to split the ownership of the world in disjoint parts.
+%% Disjoint union \oplus
+To this end, we introduce the operator $\oplus$ which is a disjoint union.
+\begin{definition}[$\oplus$, disjoint union of ownership]
+  \[
+    W_1 \oplus W_2 = W
+  \text{ iff }
+  \begin{array}[t]{l}
+    \dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand \\
+    \dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand \\
+    \dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand \\
+    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
+    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
+    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
+  \end{array}
+  \]
+\end{definition}
+
+\begin{definition}[World disjoint union $\uplus$]
+  Given worlds $W_1$, $W_2$, $W$
+  \[
+    W_1 \uplus W_2 = W
+    \text{ iff }
+    \begin{array}[t]{l}
+      \dom(\pwheap) = \dom(\pwheap[W_1]) \uplus \dom(\pwheap[W_2]) \tand \\
+      \dom(\pwfree) = \dom(\pwfree[W_1]) \uplus \dom(\pwfree[W_2]) \tand \\
+      \dom(\pwpriv) = \dom(\pwpriv[W_1]) \uplus \dom(\pwpriv[W_2]) \\
+    \end{array}
+  \]
+\end{definition}
+
+The two operators $\uplus$ and $\oplus$ are quite different.
+The difference is most clear in the treatment of pure regions: $\uplus$ allows both worlds to have the same pure region, while $\oplus$ forbids this.
+To understand this different treatment ($W_1 \uplus W_2$ and $W_1 \oplus W_2$), you should understand that the two are intended for different usages of worlds.
+The $W_1 \oplus W_2$ operator treats the worlds as specifications of authority: taking the disjoint union of worlds specifying non-exclusive ownership of a block of memory is allowed and produces a new world that also specifies non-exclusive ownership of world.
+The $W_1 \uplus W_2$  operator treats worlds as specifications of memory contents: taking the disjoint union of worlds specifying the presence of the same memory range is not allowed.
+The latter operator is used in the logical relation for components which specifies (among other things) that the world should specify the presence of the component's data memory.
+Linking two components then produces a new component with both components' data memory.
+The linked component is then valid in a world that has the combined memory presence specifications, not the combined authority.
+In other words, $\oplus$ specifies disjoint authority distribution, while $\uplus$ specifies disjoint memory allocation.
+
+Note also that this picture is further complicated by our usage of non-authority-carrying $\spatial$ regions.
+They are really only there in a world $W$ as a shadow copy of a $\spatialo$ region in another world $W'$ that $W$ will be combined with.
+The shadow copy is used for specifying when a memory satisfies a world: the memory should contain all memory ranges that anyone has authority over, not just the ones whose authority belongs to the memory itself.
+For example, if a register contains a linear pointer to a range of memory, then the register file will be valid in a world where the corresponding region is $\spatialo$, while the memory will be valid in a world with the corresponding region only $\spatial$.
+However, for the memory to satisfy the world, the block of memory needs to be there, i.e.
+the memory should contain blocks of memory satisfying every region that is $\spatialo$, $\pure$, but also just $\spatial$ (because it may be $\spatialo$ in, for example, the register file's world).
+
+%% This both makes world a specification of authority as well as a specification of memory.
+
+\subsubsection{Memory satisfaction}
+\label{subsubsec:mem-sat}
+% Well-formedness of world
+% specification satisfaction
+\[
+  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
+      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \text{ is defined} \wedge\\
+      W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
+      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}\ldotp \\
+      \quad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
+      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
+      \quad \dom(\ms_{\var{T,stack}} \uplus \var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
+      \quad \{\baddr_\stk -1,\eaddr_\stk + 1\} \in \dom(\ms_{T,f}) \wedge \\
+      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
+      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
+      \quad \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
+    \end{array}
+  \right.
+\]
+
+\[
+  \memSatStack{\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      W_\var{stack} = \pwpriv \wedge \\
+      \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
+      \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
+      \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
+      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
+      \quad \ms_0 \uplus \dots \uplus \ms_m = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World \ldotp \\
+      \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n\ldotp \\
+      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
+      \qquad \quad \exists i \ldotp \opc_i = W_\var{stack}(r).\opc \wedge \ms_i = \pi_1 (R_\ms(r))
+    \end{array}
+  \right.
+\]
+
+\[
+  \memSatFStack{ms_\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      W_\var{stack} = \pwfree \wedge \\
+      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms_\stk = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \stkb \in \dom(\ms_T) \wedge \stkb \in \dom(\ms_\stk) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
+      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}),n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
+    \end{array}
+  \right.
+\]
+
+\[
+  \lrheap(\pwheap)(W') = 
+  \left\{
+    \npair{(\overline{\sigma},ms,\ms_T)} \middle|
+    \begin{array}{l}
+      \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
+      \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
+      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
+      \quad \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
+    \end{array}
+  \right.
+\]
+
+% TODO take a look at notation and naming in this section. What do we want to call things
+\end{jversion}
 We use a type of worlds tailored to our purposes.
@@
 \subsection{The Logical Relation}
+\label{subsec:logical-relation}
 Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
@@
 
-\section{Discussion}
-\label{sec:discussion}
-% \begin{itemize}
-% \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
-% \item Sharing stack references accross component boundaries is supported
-% \item Other notions of well-bracketedness (specifically one would be to allow different stacks)
-% \end{itemize}
-\subsection{Full Abstraction}
-% - Full abstraction proofs difficult
-Our formulation of WBCF and LSE using a fully abstract overlay semantics has an important advantage with respect to others.
-Imagine that you are implementing a fully abstract compiler for a high-level language, i.e.\ a secure compiler that enforces high-level abstractions when interacting with untrusted target-language components.
-Such a compiler would need to perform many things and enforce other high-level properties than just WBCF and LSE.
-
-If such a compiler uses the \stktokens{} calling convention, then the security proof should not have to reprove security of \stktokens{}.
-Ideally, it should just combine security proofs for the compiler's other functionality with our results about \stktokens{}.
-We want to point out that our formulation enables such reuse.
-Specifically, the compiler could be factored into a part that targets \srccm{}, followed by our embedding into \trgcm{}.
-If the authors of the secure compiler can prove full abstraction of the first part (relying on WBCF and LSE in \srccm{}) and they can also prove that this first part generates well-formed and reasonable components, then full abstraction of the whole compiler follows by our result and transitivity of fully abstract compilation.
-Perhaps other reusable components of secure compilers could be formulated similarly using some form of fully abstract overlay semantics, to obtain similar reusability of their security proofs.
-
-% When creating fully-abstract compilers between low-level machines, it is a big challenge to work with the exposed addresses.
-% In particular, if the compilation changes the code size of a block of code, then it may be observable and prevent full-abstraction from being proven.
-% We would argue that when a compilation reaches a phase where addresses are exposed, then the compilation should no longer change the code.
-% This does, however, pose the challenge that there might quite a few abstractions in difference between the language where addresses are hidden to a machine where they are not.
-% We propose that this challenge is solved by implementing these abstractions in a number of overlay semantics.
-% By implementing them one by one, one can deal with one abstraction at a time reducing the complexity of each necessary full-abstraction proof.
-
-% - Need to compile to a machine with enforcement mechanisms - capability machine an option
-% - Full abstraction proofs modular, so other full abstraction proofs could target \srccm{} and thus have more abstractions to work with than if \trgcm{} was the target.
-
-%\subsection{Sharing the stack}
-% ?
-\subsection{Practical Applicability}
-We believe there are good arguments for practical applicability of \stktokens{}.
-The strong security guarantees are proven in a way that is reusable as part of a bigger proof of compiler security.
-Its costs are
-\begin{itemize}
-\item a constant and limited amount of checks on every boundary crossing.
-\item possibly a small memory overhead because stack frames must be of non-zero length
-\end{itemize}
-The main caveat is that we rely on the assumption that capability machines like CHERI can be extended with linear capabilities in an efficient way.
-
-Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team as well as our own thoughts on the matter.
-As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
-Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
-Our instructions are a bit clumsy from this point-of-view because, for example, \texttt{move} or \texttt{store} will zero the source register resp.
-memory location if the value being written is linear.
-A solution for this problem could be to add separate instructions for moving, storing and loading linear registers at the cost of additional opcode space.
-Adding splice and split will also consume some opcode space.
-
-Another problem is caused by the move semantics for \texttt{load} in the presence of multiple hardware threads.
-In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability should behave atomically, similar to a primitive compare-and-swap instruction.
-This is in principle not a problem except that atomic instructions are significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
-When using \stktokens{}, loads of linear capabilities happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, well-engineered caching could perhaps reduce the high overhead of atomic loads of linear capabilities.
-% If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
-The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
-
-\section{Related Work}
-
-In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.
-We do not repeat the work we discussed in Section~\ref{sec:introduction}.
-
-Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
-The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
-This is a recent, relatively mature capability machine which combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
-For simplicity, we have omitted features of CHERI that were not needed for \stktokens{} (e.g.\ local capabilities, virtual memory).
-
-Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
-This includes particularly the line of work on control-flow integrity~\citep{abadi_control-flow_2005}.
-This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static control graph information and a protected shadow stack.
-Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
-However, its attacker model is different, and its security goals are weaker.
-They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
-In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness saying that jumps stay within the program's static control flow graph~\cite{Abadi2005Theory}.
-Such a property ignores temporal properties and seems hard to use for reasoning.
-There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending}.
-
-More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016,juglaret_beyond_2016}.
-Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments), and we do not use per-component stacks or a trusted stack manager but a single shared stack and a decentralized calling convention based on linear capabilities.
-Both prove a secure compilation result from a high-level language which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
-
-CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
-The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.
-They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared across component boundaries.
-
-The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
-In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that applies to unsafe source languages.
-Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
-If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
-
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-\begin{acks}
-  This research was supported in part by the ModuRes Sapere Aude
-  Advanced Grant from The Danish Council for Independent Research for
-  the Natural Sciences (FNU) and by Cost Action CA15123 EUTypes.
-  Dominique Devriese held a Postdoctoral Fellowship from the Research Foundation - Flanders (FWO) during most of this research.
-\end{acks}
-
-\bibliography{references}
-
-%% Appendix
-% \appendix
-% \section{Appendix}
-
-% Text of appendix \ldots
-
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "paper"
+%%% End:

### 2019-05-14 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add more text about memory satisfaction

@@
 Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
-The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard:esop18} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.
+The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.
 
@@
 % Move to a different domain
-Therefore, we need to move to a different domain with enough structure for such an equation to have a solution.
-Specifically, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
+Therefore, we need to move to a different domain with enough structure for a solution to exist for such an equation.
+We move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
 Intuitively, a c.o.f.e. is a set with some extra structure.
@@
 In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
-We loosely follow the presentation in \citet{Birkedal:tutorial-notes}.
+We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for this particular setting.
 \begin{definition}[Ordered family of equivalences (o.f.e.)]
@@
 \end{definition}
+% TODO write something here about why the different things (the meassure of the ofe and the guaranteed limits of the cofe) gives us in terms of finding a solution.
 Functions between c.o.f.e.'s must retain the added structure relatively to sets.
@@
 However, this is not the only structure we would like to impose on our sets.
+
+% Not sure this is necessary:
+% % preordered cofe constructions
+% %% cofe to preordered cofe
+% \begin{lemma}[c.o.f.e.'s are preordered c.o.f.e's]
+%   A c.o.f.e $\cofe{X}$, is also a preordered c.o.f.e. with the order $=$.
+% \end{lemma}
+In order to express the domain equation, we must lift all parts of the world definition from sets to c.o.f.e.'s.
+That is, we need to lift the sets of partial functions, monotone non-expansive functions, and relations, respectively to c.o.f.e.'s.
+
+We need to lift the set of relations over $R$, $\Rel{R}$, to a c.o.f.e.
+However, the set of relations have no inherent step-index which means that we need to add one.
+We need to add the step-index in such a way that we can define a family of equivalences that satisfies the c.o.f.e.\ conditions.
+To this end, we define the uniform relations over $R$ $\URel{R}$.
+%% URel
+\begin{definition}[Uniform relation]
+  Given a relation $R \subseteq X \times Y$, we define $\URel{R} \subset \powerset{\nats \times R}$ as
+  \[
+    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats\ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp (n,r) \in A\}
+  \]
+  The c.o.f.e.\ $\cofe{\URel{R}}$. Define erasure as
+  \[
+    \erase{A}{n} \defeq \{(k,a) \in A \mid k < n\}
+  \]
+  and the family of equivalences as
+  \[
+    A \nequal B \text{ iff } \erase{A}{n} = \erase{B}{n}
+  \]
+\end{definition}
+The $n$-equivalences of the uniform relation basically forgets everything from the step and up and requires the rest to be equal.
+This means that everything is equal at step $0$ as everything is left out.
+
+Next, we define the c.o.f.e.\ of monotone and non-expansive functions.
+%% Monotone, non-expansive
+\begin{lemma}[The c.o.f.e. of monotone and non-expansive functions]
+  Given c.o.f.e.'s $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future_W\right)$ and $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ and a preorder $\future$ on $W$, define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ as follows
+  \[
+    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
+  \]
+\end{lemma}
+Next, we define the product c.o.f.e.
+%% Product
+\begin{lemma}[Product c.o.f.e.]
+  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \right)$
+  where the family of equivalences are defined as follows
+  \[
+    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
+  \]
+\end{lemma}
+Similarly to the product c.o.f.e. we have a sum c.o.f.e.
+%% Sum
+\begin{lemma}[Sum c.o.f.e.]
+  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \right)$ define the sum c.o.f.e.\ as $\left(X \cup Y, \left(\nequal\right)_{n=0}^{\infty}, \right)$
+  where the family of equivalences are defined as follows
+  \[
+    a \nequal b \text{ iff }
+    \left\{
+      \begin{aligned}
+        a,b \in X \wedge a \nequal_X b \vee \\
+        a,b \in Y \wedge a \nequal_Y b
+      \end{aligned}
+    \right.
+  \]
+\end{lemma}
+\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
+%% Partial function from set
+\begin{lemma}[The c.o.f.e.\ of partial functions]
+  Given a set $S$ and a c.o.f.e $P$, the set of partial non-expansive functions $S \parfun P$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
+  \[
+    f \nequal[0] g \wedge f \nequal g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \nequal g(x)
+  \]
+\end{lemma}
+
 In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
@@
 \end{lemma}
-% Not sure this is necessary:
-% % preordered cofe constructions
-% %% cofe to preordered cofe
-% \begin{lemma}[c.o.f.e.'s are preordered c.o.f.e's]
-%   A c.o.f.e $\cofe{X}$, is also a preordered c.o.f.e. with the order $=$.
-% \end{lemma}
-In order to express the domain equation, we must lift all parts of the world definition from sets to c.o.f.e.'s.
-That is, we need to lift the sets of partial functions, monotone non-expansive functions, and relations, respectively to c.o.f.e.'s.
-
-We need to lift the set of relations over $R$, $\Rel{R}$, to a c.o.f.e.
-However, the set of relations have no inherent step-index which means that we need to add one.
-We need to add the step-index in such a way that we can define a family of equivalences that satisfies the c.o.f.e.\ conditions.
-To this end, we define the uniform relations over $R$ $\URel{R}$.
-%% URel
-\begin{definition}[Uniform relation]
-  Given a relation $R \subseteq X \times Y$, we define $\URel{R} \subset \powerset{\nats \times R}$ as
-  \[
-    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats\ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp (n,r) \in A\}
-  \]
-  The c.o.f.e.\ $\cofe{\URel{R}}$. Define erasure as
-  \[
-    \erase{A}{n} \defeq \{(k,a) \in A \mid k < n\}
-  \]
-  and the family of equivalences as
-  \[
-    A \nequal B \text{ iff } \erase{A}{n} = \erase{B}{n}
-  \]
-\end{definition}
-The $n$-equivalences of the uniform relation basically forgets everything from the step and up and requires the rest to be equal.
-This means that everything is equal at step $0$ as everything is left out.
-
-Next, we define the preordered c.o.f.e.\ of monotone and non-expansive functions.
-%% Monotone, non-expansive
-\begin{lemma}[Preordered c.o.f.e. of monotone and non-expansive functions]
-  Given preordered c.o.f.e.\ $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future_W\right)$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ as follows
-  \[
-    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
-  \]
-  and preorder
-  \[
-    h \future h' \text{ iff } \forall r \in \dom(h) \ldotp h(w) = h'(w)
-  \]
-\end{lemma}
-Next, we define the product preordered c.o.f.e.
-%% Product
-\begin{lemma}[Product preordered c.o.f.e.]
-  Two preordered c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ can be composed to a product preordered c.o.f.e $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
-  where the family of equivalences are defined as follows
-  \[
-    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
-  \]
-  and the order as follows
-  \[
-    (x,y) \future (x',y') \text{ iff } x \future_X x' \wedge y \future_Y y'
-  \]
-\end{lemma}
-We use the product preordered c.o.f.e.\ when we just need to compose two existing preordered c.o.f.e.'s into a new preordered c.o.f.e.
-However, in the case of our regions, this do not suffice as they need to use the future region relation we defined in Section~\ref{subsubsec:ft-and-revocation}.
-We therefore define separate preordered c.o.f.e.'s that we use to define these.
-\begin{lemma}[$\Regions$ preordered c.o.f.e.]
-  Given preordered c.o.f.e.\ $\left(H, \left(\nequal_H\right)_{n=0}^{\infty}, \future_H \right)$ define $\left(\{\spatial\} \times H + \{\spatialo\} \times H + \revoked, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
-  where $\future$ is the future region relation from Section~\ref{subsubsec:ft-and-revocation} and
-  \[
-    r \nequal r' \text{ iff } n=0 \vee \left\{
-      \begin{aligned}
-      (\_,h),(\_,h') \in \{ \spatial  \} \times H \wedge h \nequal h' \vee \\
-      (\_,h),(\_,h') \in \{ \spatialo \} \times H \wedge h \nequal h' \vee \\
-      r,r' = \{revoked\}      
-    \end{aligned}\right.
-  \]
-\end{lemma}
-
-%% Sum
-Sum c.o.f.e.'s are defined in a similar manner.
-
-\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
-%% Partial function from set
-\begin{lemma}[Preordered c.o.f.e.\ of partial functions]
-  Given a set $S$ and a preorder c.o.f.e $P$, the set of partial functions $S \parfun P$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
-  \[
-    f \nequal[0] g \wedge f \nequal g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \nequal g(x)
-  \]
-  and preorder
-  \[
-    f \future g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \future g(x)
-  \]
-\end{lemma}
-
+% TODO a preordered c.o.f.e. w/ our world?
+
+In some sense, the step-index in a c.o.f.e. gives us a metric that we may use to find a fixed-point.
+However, for a fixed-point to exist, the metric must decrease before a self-reference.
+As this is not an inherent property of c.o.f.e.'s, we have a c.o.f.e.\ construction called \emph{later} with the sole purpose of decrementing the index.
 %% Later cofe
@@
   \[
-    x \nequal[0]' x' \wedge x \nequal[n]' x' \texttt{ iff } x \nequal[n+1] x'
+    x \nequal[0]' x' \wedge x \nequal[n]' x' \text{ iff } x \nequal[n+1] x'
   \]
 \end{lemma}
-
-% partial functions
-%
-
-% New world definitions
-The world definitions we presented so far needs to move to the domain of preordered c.o.f.e's.
-Morally, the definitions are the same, though.
-The definition of the two kinds of regions now uses $\URel{}$ rather than $\Rel$, and the world indexed-functions are monotone.
-Finally, all functions are now non-expansive.
+The later needs to guard all negative self-references in a recursive domain equation in order for a solution to exist.
+The modality does not have to be placed directly on the self-reference.
+For instance, in the following it is placed on the world itself.
+The later placement does, however, affect what the logical relation definitions in Section~\ref{subsec:logical-relation} looks like, so in the end it is a matter of taste.
+% Solution theorem
+\begin{theorem}
+  \label{thm:rec-dom-eq-sol}
+  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
+  \[
+    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
+  \]
+  and for $\hat{W},\hat{W}' \in \Wor$
+  \[
+    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
+  \]
+  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
+\[
+  \Worldh = \RegionName \parfun (\Regions + \Regionh)
+\]
+and
+\[
+  \Worlds = \RegionName \parfun (\Regions \times \Addr)
+\]
+and
+\[
+  \Worldfs = \RegionName \parfun \Regions
+\]
+where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
 \begin{multline*}
@@
   \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
-  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps^2})
 \end{multline*}
@@
 \]
-
-% Reference to solution?
+\end{theorem}
+% TODO informally describe how solution to recursive domain equation is found (in terms of what we have to show, not the fixed-point theorem).
+%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
 %% Aleš note? Tutorial note?
-%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
-
-% Argue, informally, that it is okay
-
-% Solution theorem
+
+The domains in Theorem~\ref{thm:rec-dom-eq-sol} are made from the c.o.f.e.\ and preorder c.o.f.e.\ constructions we presented earlier in this section.
+They are also the definitions we use in our world definition.
+We do not define $\Wor$ to be our worlds.
+Instead, we define our world as
+\[
+  \World = \Worldh \times \Worlds \times \Worldfs
+\]
+which is easier as we only have to use $\xi$ and move under a later when we have gone into the region and need to work with $\Wor$.
 
 \subsubsection{Joining worlds}
-
+% TODO rewrite this introduction
 In order to express this, we will need to split the ownership of the world in disjoint parts.
@@
 \label{subsubsec:mem-sat}
-% Well-formedness of world
-% specification satisfaction
-\[
-  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
+% Specification
+The world can be seen as a specification of the memory contents.
+This means that we need to define what it means for a pair of \trgcm{} and \srccm{} memories to satisfy the specification.
+% Call stack
+The world also keeps track of the structure of the call stack, the allowed uses of designated seals, and linear capability authority, so these things also influence memory satisfaction.
+% Well-formedness
+The memory satisfaction also acts as a well-formedness judgement.
+Our world definition allows regions to be overlapping in the sense that two regions may allow memory segments with the same addresses.
+We want the regions to be disjoint, so we easily can get the invariants of each memory segment.
+
+Memory satisfaction is split into four definitions.
+At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory, $\ms_S$, $\ms_\stk$, and $\stk$, from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
+The top memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
+It also splits the world in three to distribute the authority of the world.
+The three parts of $\ms_T$ are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrheap$, $\lrstk[]$, and $\lrfree[]$, respectively. 
+The $\lrheap$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
+The $\lrheap$ relation also ensures that the seal invariants associated with each regions do not put invariants on the same seals.
+The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
+The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the stack frames have the correct order relative to each other as well as the base address of the stack.
+Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
+$\lrfree[]$ also makes sure that that base address of the stack is actually part of the stack.
+\begin{definition}[Heap relation]
+For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrheap$ as:
+\[
+  \npair{(\overline{\sigma},ms,\ms_T)} \in \lrheap(\pwheap)(W') = 
   \left\{
     \begin{array}{l}
-      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
-      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \text{ is defined} \wedge\\
-      W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
-      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}\ldotp \\
-      \quad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
-      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
-      \quad \dom(\ms_{\var{T,stack}} \uplus \var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
-      \quad \{\baddr_\stk -1,\eaddr_\stk + 1\} \in \dom(\ms_{T,f}) \wedge \\
-      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
-      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
-      \quad \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
+      \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
+      \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
+      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
+      \quad \forall r \in \dom(\activeReg{\pwheap})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
     \end{array}
@@
 \]
-
+\end{definition}
+Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
+The $\lrheap$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.
+
+To a large extend, the way $\lrheap$ is defined is pretty standard.
+$\lrheap$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into pairs each satisfying the invariant of a region.
+The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction also partition the authority of the world.
+Each of the memory segment pairs must be in the region invariant with respect to a specific world partition which makes sure that uniqueness of linearity of capabilities is respected.
+
+The heap sub-world contains all seal invariants.
+Similar to memory segments, only one seal invariant should impose restrictions on a seal which $\lrheap$ makes sure is the case.
+\begin{definition}[Free stack relation]
+\[
+  \memSatFStack{ms_\stk,\ms_T}{W} \text{ iff } 
+  \left\{
+    \begin{array}{l}
+      \gc = (\_,\stkb,\_,\_) \wedge\\
+      W_\var{stack} = \pwfree \wedge \\
+      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \MemSeg \wedge \\
+      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_2(R_\ms(r)) \wedge \\
+      \quad \ms_\stk = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \stkb \in \dom(\ms_T) \wedge \stkb \in \dom(\ms_\stk) \wedge \\
+      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
+      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
+      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}),n' < n \ldotp \\
+      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
+    \end{array}
+  \right.
+\]
+\end{definition}
+The free stack relation $\lrfree[]$ is in most regards like the heap relation.
+The free stack relation, partitions the \srccm{} and \trccm free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under the
+\begin{definition}[Stack relation]
 \[
@@
     \begin{array}{l}
+      \gc = (\_,\stkb,\_,\_) \wedge \\
       W_\var{stack} = \pwpriv \wedge \\
@@
 \]
-
-\[
-  \memSatFStack{ms_\stk,\ms_T}{W} \text{ iff } 
+\end{definition}
+The stack relation $\lrstk[]$ is similar to the heap relation in some ways.
+$\lrstk[]$ also partitions the memory, but it only partitions the \trgcm{} as the partitioning of the \srccm{} is already given as the stack frames.
+
+\begin{definition}[Memory satisfaction]
+  For memory segments $\ms_S$, $\ms_\stk$, and $\ms_T$, stack $\stk$, and world $W$ we define memory satisfaction as
+\[
+  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
   \left\{
     \begin{array}{l}
-      W_\var{stack} = \pwfree \wedge \\
-      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \MemSeg \wedge \\
-      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_2(R_\ms(r)) \wedge \\
-      \quad \ms_\stk = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
-      \quad \stkb \in \dom(\ms_T) \wedge \stkb \in \dom(\ms_\stk) \wedge \\
-      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
-      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
-      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}),n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
+      \exists \opc_0, \dots \opc_m, \ms_0, \dots, \ms_m, W_{\var{stack}}, W_{\var{free\_stack}}, W_{\var{heap}} \ldotp\\
+      \quad \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
+      \quad \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \text{ is defined} \wedge\\
+      \quad W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
+      \quad\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}\ldotp \\
+      \qquad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
+      \qquad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
+      \qquad \dom(\ms_{\var{T,stack}} \uplus \var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
+      \qquad \{\baddr_\stk -1,\eaddr_\stk + 1\} \in \dom(\ms_{T,f}) \wedge \\
+      \qquad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
+      \qquad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
+      \qquad \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
     \end{array}
@@
 \]
-
-\[
-  \lrheap(\pwheap)(W') = 
-  \left\{
-    \npair{(\overline{\sigma},ms,\ms_T)} \middle|
-    \begin{array}{l}
-      \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
-      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
-      \quad \ms = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
-      \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
-      \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
-      \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
-      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
-      \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
-      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
-      \quad \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
-    \end{array}
-  \right.
-\]
+\end{definition}
 

### 2019-05-14 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Remove the old world section

@@
 \end{jversion}
-We use a type of worlds tailored to our purposes.
-They consist of three sub-worlds: $\Wor = \Worldh \times \Worlds \times \Worldfs$, capturing assumptions about the heap\footnote{Actually, we use the term heap to describe all memory except the stack, including, for example, code memory.}, the inactive and the active part of the stack, respectively.
-Sub-worlds consist of a finite mapping from region names to regions which come in two forms (spatial and shared):
-\begin{align*}
-  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
-  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
-  \Worldfs &={} \RegionName \parfun \Regions
-\end{align*}
-Different parts of the world can contain different types of regions: heap-related assumptions can be either spatial or shared (see below), while stack-related regions must be spatial.
-Additionally, regions for inactive parts of the stack additionally include an address specifying the return address for that stack frame.
-
-Regions in $\Regionh$ specify the presence of an invariant in the system, shared with the rest of the system.
-They are tagged with the syntactic token $\pure$ and may prescribe two different types of requirements:
-\begin{align*}
-  \Regionh &= 
-             \left\{\begin{multlined}
-               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
-               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
-             \end{multlined}\right.
-\end{align*}
-First, they may require the presence of \srccm{} and \trgcm{} memory segments satisfying a given relation in $\Wor \monnefun \URel{\MemSeg^2}$ (readers unfamiliar with Kripke and step-indexed logical relations may read $\Wor \monnefun \URel{\MemSeg^2}$ as the set of functions from $\Wor$ to relations on memory segments).
-A region might, for example, require the presence of a certain list of instructions at a certain set of memory addresses in both \srccm{} and \trgcm{}.
-If a memory segment is owned by a given region, then that memory segment must be disjoint from the memory segments owned by all other regions.
-Second, a shared region may also contain a partial function from seals to relations on pairs of sealable capabilities: $\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps}$.
-When the region defines such a relation for a given seal, then no other region in the world can do the same, and any value signed with that seal will be required to satisfy the registered relation.
-
-Spatial regions are similar to shared regions, but they are tagged as $\spatial$ or $\spatialo$ and may not specify seal invariants.
-Additionally, a spatial region may also be $\revoked$:
-\begin{align*}
-  \Regions &={}
-             \begin{multlined}
-               \{\spatial,\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \{\revoked\}
-             \end{multlined}
-\end{align*}
-
-The difference between spatial and shared regions is related to linearity and ownership.
-For example, a \trgcm{} linear capability to a piece of memory is related to its \srccm{} counterpart, but only if no other linear capability overlaps with it.
-We will model such an assumption of exclusive ownership by making the relatedness rely on the presence of a $\spatialo$ region that only one value in the system may rely on.
-More concretely, we will define how to combine worlds $W_1$ and $W_2$ into a combined world $W_1 \oplus W_2$, on the condition that they represent compatible assumptions: $W_1$ and $W_2$ must contain the same regions except that they must respect exclusive ownership: $\pure$ regions must be present in both worlds, but a $\spatialo$ region can only be present in one and must be $\spatial$ in the other.
-When $W_1 \oplus W_2$ is defined, we say that worlds $W_1$ and $W_2$ are compatible and we refer to $W_1$ and $W_2$ as compatible partitions of the combined world.
-
-The reason that we have $\spatial$ regions (in addition to $\spatialo$ ones) is for defining when \srccm{} and \trgcm{} memories are related.
-We need them to contain suitable memory segments for all regions in the system, even for regions owned by values that live outside the memory (for example register values).
-Those regions will be in the world for the memory, but only as $\spatial$, i.e.\ they may not actually be referenced from within memory, but we still require them to be backed by suitable memory contents.
-We also use a relation $W_2 \future W_1$ that defines when a world $W_2$ is a future world of a world $W_1$.
-Relations that hold with respect to $W_1$, will then generally continue to hold in $W_2$.
-Our future world relation is fairly standard: the future world must contain all the previous world's regions, except that $\spatial$ regions are allowed to become $\spatialo$ (i.e.\ gaining ownership of a region will never break relatedness) or $\revoked$ (i.e.\ revoking spatial regions will never break relatedness).
-
-Attentive readers may have noticed that our definition of worlds is actually cyclic: worlds in $\Wor$ contain regions in $\Regionh$, but those contain partial functions from the set $\Wor$ to something else.
-Such recursive worlds are in fact common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} (essentially an advanced form of step-indexing) to construct the set $\Wor$ and rigorously resolve the circularity.
-% \begin{lemma}
-%   \label{thm:recursive-domain-eq}
-%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
-%   \[
-%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
-%   \]
-%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
-% \end{lemma}
-
-In our proof, we use only a few different types of regions.
-For space reasons, we do not go into their definitions (see the \citet{technical_report}), but we give a brief overview here.
-First, we have the code region $\codereg{\sigrets,\sigcloss,\mscode}$.
-This is a shared region that represents the assumption that memory segment
-$\mscode$ is loaded in heap memory at a certain location, it is well-formed and it uses return seals $\sigrets$ and closure seals $\sigcloss$.
-% Lau: I cannot parse this sentence:
-It takes ownership of those seals and registers appropriate invariants on the capabilities that may be signed with them.
-
-A second type of regions $\stdreg{A}{\pure}$ and $\stdreg{A}{\spatialo}$ governs heap or stack memory at a set of addresses $A$.
-It simply requires the presence of a memory segment for those addresses, such that the memory contains values that themselves satisfy the relation between values that we will see below.
-The contents of the memory is allowed to change as long as the new contents are still valid.
-Finally, a third type of regions $\stareg[(\ms_S,\ms_T)]{\spatialo}$ requires the presence of two given memory segment $\ms_S$ and $\ms_T$ and does not allow them to change.
-This region is used, for example, to govern inactive parts of the stack whose contents is required to remain unmodified.
 

### 2019-05-14 popl-2019-journal-v/proving-full-abstraction.tex
Fix small typo in POPL technical report
[POPLjv] Add labels to memory satisfaction definitions

@@
 \begin{definition}[Heap relation]
+\label{def:heap-rel}
 For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrheap$ as:
@@
 \begin{definition}[Free stack relation]
+\label{def:free-stack-rel}
 \[
@@
 \begin{definition}[Stack relation]
+  \label{def:stack-rel}
 \[
@@
       \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
-      \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
+      \quad \forall j < i \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
       \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\

### 2019-05-14 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Finish first draft of mem sat section

@@
 \subsubsection{Joining worlds}
-% TODO rewrite this introduction
+% TODO rewrite this introduction, add text between definitions (?), and reread this section.
 In order to express this, we will need to split the ownership of the world in disjoint parts.
@@
 The free stack relation $\lrfree[]$ is in most regards like the heap relation.
-The free stack relation, partitions the \srccm{} and \trccm free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under the
+The free stack relation, partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under the
 \begin{definition}[Stack relation]
@@
 The stack relation $\lrstk[]$ is similar to the heap relation in some ways.
-$\lrstk[]$ also partitions the memory, but it only partitions the \trgcm{} as the partitioning of the \srccm{} is already given as the stack frames.
-
+$\lrstk[]$ also partitions the memory but only the \trgcm{} as the \srccm{} partitions are given as the stack frames, and it also partitions the authority of the world, so it can relate the stack frames such that they respect linearity.
+The stack on \srccm{} represents the call stack which means that each stack frame represents a call and its local data.
+The operational semantics of \trgcm{} does not enforce the same stack structure, so we make sure that a \srccm{} stack is only related to \trgcm{} stack that has the same structure.
+That is for a memory segment that represents a stack frame, all the addresses of memory frames lower in the stack should have strictly smaller memory addresses .
+Further, the stack frames should be in the part of the memory we agree to be the stack which means that the addresses should be less than $\stkb$.
+Informally, this just means that the stack should be laid out in memory as a downwards growing stack with no addresses below $\stkb$.
+% TODO add figure that illustrates this.
+According to \stktokens{} (described in \sectionname~\ref{sec:stktokens-explained}), stack frames must be non-empty, so they are distinguishable from a missing stack frame.
+For this reason, $\lrstk[]$ requires every stack frame to be non-empty.
+Each stack frame in the \srccm{} stack contains an old program pointer.
+Each of the old program counters on the stack should correspond to the old program counter recorded in the region of the world associated with the stack frame.
+To achieve this, the partitioning function $R_\ms$ also records an $\opc$ for each region, and this $\opc$ should establish the link the region and the stack frame.
+
+In order to tie Definition~\ref{def:heap-rel},~\ref{def:free-stack-rel},~and~\ref{def:stack-rel}, we define memory satisfaction.
+Memory satisfaction defines when a \srccm{} memory , consisting of a heap, a stack, and a free stack, relates to a \trgcm{} memory under a world.
 \begin{definition}[Memory satisfaction]
@@
       \qquad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
-      \qquad \dom(\ms_{\var{T,stack}} \uplus \var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
-      \qquad \{\baddr_\stk -1,\eaddr_\stk + 1\} \in \dom(\ms_{T,f}) \wedge \\
+      \qquad \dom(\ms_{\var{T,stack}} \uplus \ms_\var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
+      \qquad \baddr_\stk -1,\eaddr_\stk + 1 \in \dom(\ms_{T,f}) \wedge \\
       \qquad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
@@
 \end{definition}
+Memory satisfaction partitions the \trgcm{} memory in a heap, stack frames, a free stack and a frame.
+The \srccm{} heap is split in two: the active heap and a frame.
+Our configurations describe the complete machine state, but we may only be interested in the invariants on part of it.
+The frame allows us to basically ignore the part of the memory that won't affect the computation.
+Just like the previous memory relations, the world is split in three to make sure that linearity is respected.
+Each part of the \srccm{} memory is related to the appropriate part of the memory from \trgcm{} by the relevant relation under a partition of the world.
+
+\stktokens{} requires the stack to not be adjacent to heap or code memory.
+This is enforced in the memory satisfaction by requiring that the addresses next to the memory are in the frame.
 
 % TODO take a look at notation and naming in this section. What do we want to call things
+% TODO look over the original POPL paper to make sure to reuse all the text that makes sense to reuse.
+% TODO add an introduction to c.o.f.e.'s that describe how standard c.o.f.e. constructions always can be used to construct something with a solution.
 \end{jversion}

### 2019-05-14 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add LR definitions from TR

@@
 These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another use of step-indexing.
-For space reasons, we cannot show all of these definitions, but we will try to give an overview.
-
+In Section~\ref{subsubsec:mem-sat}, we already defined memory satisfaction, the relation for memories.
+In the following, we define each of the remaining relations and give some intuition about the definitions.
+The logical relation we define ends up as a recursive definition.
+The circularity is resolved by a step-index in the definitions, but the circularity also poses a chicken and egg kinda situation with respect to the order in which the definitions of the relations should be presented.
+There is no canonical way of presenting the logical relation as we are bound to make forward references.
+For this reason, we suggest making a cursorily first read through followed by a more thorough read.
+
+\subsubsection{Value Relations}
+% TODO rewrite the below text
+% Should in general terms describe value relation
+% Maybe have a figure that sketches it like now.
 Note first how we have two value relations, whose definitions are sketched in Figure~\ref{fig:value-relation}.
@@
 
-Details are in the \cite{technical_report}, but roughly, for stack capabilities, the omitted condition requires that the world contains a $\spatialo$ region governing this part of the stack.
+Roughly, for stack capabilities, the omitted condition requires that the world contains a $\spatialo$ region governing this part of the stack.
 For memory capabilities $((\perm,\lin),\baddr,\eaddr,\aaddr)$, a region in the world must govern memory $[\baddr,\eaddr]$, either $\spatialo$ or $\pure$, depending on the linearity $\lin$ of the capability.
@@
 
-\begin{figure}
-  \centering
+% \begin{figure}
+%   \centering
+% \begin{align*}
+%   \lrrg{\trust}(W) &= \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+%     \begin{array}{l}
+%     % Lau: Consider using \Wor instead of \World as we have not made a distinction between the two.
+%       \exists S : (\RegName \setminus \{\pcreg \})\fun \World \ldotp \\
+%       \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg \} \cup R))} S(r) \wedge \\
+%       \quad \forall r \in \RegName \setminus \{\pcreg \}\ldotp \npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
+%     \end{array}
+%             \right\}\\
+%   \lre(W)&= \left\{ \begin{array}{l}
+%     \npair{\stpair[.]{w_{c,S}}{w_{c,T}}} | \\
+%     \quad\forall \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
+%     \qquad\npair{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
+%     \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{w_{c,S}} \tand\\
+%                      \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{w_{c,T}} \tand \\
+%                      \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+%     \qquad\qquad\Rightarrow\npair{\left(\Phi_S', \Phi_T' \right)}\in \lro
+%   \end{array}
+%   \right\}
+% \end{align*}
+%   \begin{align*}
+%   \lro[] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
+%     \src{\Phi_S \sterm{}} \Leftrightarrow \Phi_T \trg{\term} \}
+% \end{align*}
+% \caption{Simplified sketches of the register file relation $\lrr(W)$, the relation
+%   for $\com{jmp}$ targets $\lre(W)$  and the observation relation $\lro(W)$.}
+% \label{fig:obs-rel}
+% \end{figure}
+
+\paragraph{Permission based conditions}
+\[
+  \addressable{\lin,W} =
+  \begin{cases}
+    \{ r \mid W(r) = (\pure,\_) \} & \text{if $\lin = \normal$} \\
+    \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
+  \end{cases}
+\]
+
+\[
+  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
+    \end{array}
+  \right\}
+\]
+\[
+  \stackReadCond{W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
+      \qquad \forall r \in S \ldotp |R(r)| = 1\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Section~\ref{par:standard-regions}.
+\[
+  \xReadCond{W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\normal, \pwheap} \ldotp \\
+      \qquad \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\wedge\\
+      \qquad \dom(\code) \supseteq A 
+    \end{array}
+  \right\}
+\]
+
+\begin{definition}
+  \label{def:address-stratified}
+  We say that a region $\iota = (\_,H,\_)$ is address stratified iff
+  \[
+    \begin{array}{l}
+      \forall n, \src{\ms_S},\ms_T,\src{\ms_S'},\ms_T',s,\hat{W}\ldotp \\
+      \quad \npair{\stpair{\ms}{\ms}}, \npair{\stpair[.]{\ms_S'}{\ms_T'}} \in H \; \hat{W} \wedge \\
+      \quad \dom(\src{\ms_S}) = \dom(\ms_T) = \dom(\src{\ms_S'}) = \dom(\ms_T') \\
+      \quad \Rightarrow \\
+      \qquad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\src{\ms_S}\update{\aaddr}{\src{\ms_S'}(\aaddr)},\ms_T\update{\aaddr}{\ms_T'(\aaddr)})} \in H \; \hat{W}
+    \end{array}
+  \]
+\end{definition}
+\[
+  \writeCond{\lin,W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
+      \qquad \quad \pwheap(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Paragraph~\ref{par:standard-regions}.
+\[
+  \stackWriteCond{W} = \left\{ \npair{A}) \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
+      \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
+      \qquad \quad \pwfree(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\]
+where $\iota_A$ is a standard region defined in Paragraph~\ref{par:standard-regions}.
 \begin{align*}
-  \lrrg{\trust}(W) &= \left\{ \npair{\stpair{\reg}{\reg}} \middle|
-    \begin{array}{l}
-    % Lau: Consider using \Wor instead of \World as we have not made a distinction between the two.
-      \exists S : (\RegName \setminus \{\pcreg \})\fun \World \ldotp \\
-      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg \} \cup R))} S(r) \wedge \\
-      \quad \forall r \in \RegName \setminus \{\pcreg \}\ldotp \npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
+  \execCond{W} &=
+  \left\{ \npair{A} \middle|
+    \begin{array}{l}
+      \forall n' < n, W' \future \purePart{W}, \aaddr \in [\baddr',\eaddr'] \subseteq A \ldotp\\
+      \quad \npair[n']{\stpair[.]{((\rx,\normal),\baddr',\eaddr',\aaddr)}{((\rx,\normal),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+    \end{array}
+    \right\}
+\end{align*}
+
+\paragraph{Standard regions}
+\label{par:standard-regions}
+Standard region:
+\[
+  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc) , v \in \{\spa,\spao\}
+\]
+for readability, we use $\spao$ short for $\spatialo$, $\spa$ short for $\spatial$, and $\pur$ as short for $\pure$.
+\[
+  \stdreg{A,\gc}{\pur} \defeq (\pur,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
+\]
+where $H^{\mathrm{std},\square}_A$ is defined as follows:
+\[
+  H_A^{\mathrm{std},\square} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
+    \begin{array}{l}
+      \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
+      \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
+      \quad \forall \aaddr \in A \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+    \end{array}
+  \right\}
+\]
+
+\[
+  \stareg[\stpair{\ms}{\ms},\gc]{v,\square} = (v,H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc)
+\]
+\[
+  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+    \begin{array}{l}
+      \exists S : \dom(\ms) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
+      \quad \forall \aaddr \in \dom(\ms) \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+    \end{array}
+\right\}
+\]
+
+\[
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
+H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc,
+H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+\]
+
+\begin{multline*}
+  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
+  \left\{\npair{\arraycolsep=0pt\left(\array{l}\code \uplus \mspad,\\ \code \uplus \mspad\endarray\right)} \middle|
+    \begin{array}{l}
+    \dom(\code) = [\baddr,\eaddr] \wedge \\
+      ([\baddr - 1, \eaddr + 1] \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
+      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
+      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
+      \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \code \wedge\\
+      \forall a \in \dom(\code)\ldotp\\
+      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
+    \end{array}
+  \right\}
+\end{multline*}
+
+\begin{multline*}
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  \begin{array}[t]{l}
+\left\{
+    \begin{array}{l}
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+      \begin{array}{l}
+        \sigrets \subseteq \gsigrets \tand \\
+        \dom(\code) \subseteq \ta \tand\\
+        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \aaddr = \aaddr' + \retoffset \tand \\
+        \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
+        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
+      \end{array}
+    \end{array}
+      \right\} \uplus \\
+\left\{
+    \begin{array}{l}
+\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
+      \begin{array}{l}
+        \sigrets \subseteq \gsigrets \tand \\
+        \dom(\code) \subseteq \ta \tand\\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
+        \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
+        \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
+      \end{array}
+    \end{array}
+    \right\} \\
+  \end{array}\\
+    \text{ if } \sigma \in \sigrets\\
+\end{multline*}
+and
+\begin{multline*}
+  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  \left\{
+    \begin{array}{l}
+\left. \npair{(\vsc, \vsc' )} \middle| \right. \\
+      \begin{array}{l}
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
+        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
+         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
+         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
+      \end{array}
     \end{array}
   \right\}\\
-  \lre(W)&= \left\{ \begin{array}{l}
-    \npair{\stpair[.]{w_{c,S}}{w_{c,T}}} | \\
-    \quad\forall \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
-    \qquad\npair{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
-    \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{w_{c,S}} \tand\\
-                     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{w_{c,T}} \tand \\
-                     \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-    \qquad\qquad\Rightarrow\npair{\left(\Phi_S', \Phi_T' \right)}\in \lro
-  \end{array}
-  \right\}
-\end{align*}
-  \begin{align*}
-  \lro[] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
-    \src{\Phi_S \sterm{}} \Leftrightarrow \Phi_T \trg{\term} \}
-\end{align*}
-\caption{Simplified sketches of the register file relation $\lrr(W)$, the relation
-  for $\com{jmp}$ targets $\lre(W)$  and the observation relation $\lro(W)$.}
-\label{fig:obs-rel}
-\end{figure}
-
+  \text{ if } \sigma \in \sigcloss
+\end{multline*}
+
+\paragraph{The untrusted value relation}
+\[
+  \lrv(W) =
+  \begin{array}[t]{l}
+    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
+%
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
+    \begin{array}{l}
+      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
+      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
+      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
+      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
+      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
+
+      % This case was never used (?) so it is not necessary.
+      % \wedge \\
+      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
+    \end{array}
+    \right\}\cup\\
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
+    \; \middle| \;
+    \begin{array}{l}
+      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
+      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+    \end{array}
+    \right\} \cup \\
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \begin{array}{l}
+      \begin{array}{r l l}
+        \perm \not\in \{\rx,\rwx\} \wedge\\
+        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
+      \end{array}
+    \end{array}
+    \right\} \cup \\
+%
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \begin{array}{l}
+      [b,e] \mathrel{\#} \ta \tand\\
+      \begin{array}{r l l }
+        % \perm = \noperm & \Rightarrow & \npair{(\lin,\baddr,\eaddr)} \in \lrp(W) \wedge \\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \readCond{\lin,W} \wedge\\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \writeCond{\lin,W} \wedge\\
+        % we are excluding rwx pointers.
+        \perm \neq \rwx \wedge \\
+        % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
+        %                            \npair{(\baddr,\eaddr)} \in \xReadCond{\lin,W} \endarray\\
+        \perm = \rx &\Rightarrow& \array[t]{l}\npair{[\baddr,\eaddr]} \in \execCond{W} \wedge\\
+        \npair{[\baddr,\eaddr]} \in \xReadCond{W} \wedge \\
+                                  \lin = \normal \\ \endarray
+      \end{array}
+    \end{array}
+    \right\}
+  \end{array}
+\]
+
+\paragraph{The trusted value relation}
+\[
+  \lrvtrusted[\square,\gc](W) =
+  \begin{array}[t]{l}
+    \lrv(W)\cup \\
+%
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
+    \; \middle| \;
+    \begin{array}{l}
+      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
+      \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)} \wedge\\
+      \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
+      \quad \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss
+    \end{array}
+    \right\} \cup \\
+    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
+    \begin{array}{l}
+      \perm \sqsubseteq \rx \wedge \\
+      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
+      {} [\baddr,\eaddr] \subseteq \ta \wedge\\
+      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
+    \end{array}
+    \right\}
+  \end{array}
+\]
+
+\subsubsection{Register file relation}
 In Figure~\ref{fig:obs-rel}, we show sketches of the register file relation $\lrr$, the relation between $\tjmp{}$ targets $\lre(W)$ and the observation relation $\lro$.
@@
 The $\lre(W)$ relation then includes any two words which can be plugged into related register files and memories (in compatible worlds), to obtain execution configurations in the observation relation.
+
+\[
+  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+    \begin{array}{l}
+      \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
+      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \wedge \\
+      \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
+      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
+    \end{array}
+            \right\}
+\]
+We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
+
+\begin{lemma}[Untrusted is trusted]
+  \label{lem:untrusted-supset-trust}
+  \begin{itemize}
+  \item $\lrvtrusted(W) \supseteq \lrv(W)$
+  \item $\lrrtrusted(W) \supseteq \lrr(W)$
+  \end{itemize}
+\end{lemma}
+
+\subsubsection{Observation relation}
+\[
+  \lro[\preceq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \forall i \leq n \ldotp \\
+      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
+    \end{array}
+\right\}
+\]
+\[
+  \lro[\succeq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \forall i \leq n \ldotp \\ 
+      \quad (\ms_T ,\reg_T) \term[i] \\\qquad\Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
+    \end{array}
+\right\}
+\]
+
+\subsubsection{Expression relations}
+% Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
+% The argument for having one relation relate pairs of pairs of capabilities and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
+\begin{align*}
+  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+    \begin{array}{l}
+      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad \forall W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
+      \qquad\Phi_T = (\ms_T,\reg_T) \wedge\\
+      \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
+      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp\\
+      \quad\qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+      \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\\
+      \quad\qquad\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}\\
+  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
+    \begin{array}{l}
+      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad \forall W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
+      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
+      \qquad\Phi_T = (\ms_T,\reg_T)\\
+      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
+      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}
+\end{align*}
 

### 2019-05-15 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Move observation relation section
[POPLjv] Write value relation introduction

@@
 \subsection{The Logical Relation}
+\begin{jversion}
 \label{subsec:logical-relation}
@@
 
+\subsubsection{Observation relation}
+% TODO write section
+\[
+  \lro[\preceq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \forall i \leq n \ldotp \\
+      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
+    \end{array}
+\right\}
+\]
+\[
+  \lro[\succeq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \forall i \leq n \ldotp \\ 
+      \quad (\ms_T ,\reg_T) \term[i] \\\qquad\Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
+    \end{array}
+\right\}
+\]
+
 \subsubsection{Value Relations}
-% TODO rewrite the below text
-% Should in general terms describe value relation
-% Maybe have a figure that sketches it like now.
-Note first how we have two value relations, whose definitions are sketched in Figure~\ref{fig:value-relation}.
-The difference is that the untrusted value relation $\lrv(W)$ does not just express that the two values are related, but also that they are safe to pass to an untrusted adversary, i.e.
+The value relation relates \trgcm{} words to \srccm{} words.
+The \srccm{} machine has special tokens that represent the stack capabilities and the return pointer components.
+These tokens do not exist on \trgcm{}, but all of the tokens correspond to capabilities on \trgcm{}, and the value relation establishes the link between then.
+\citet{skorstengaard_reasoning_2017} defines a logical relation that can be seen as a notion of capability safety.
+When they define their value relation, they define based on the question ``What is the most we can allow an adversary to do with this word without breaking memory guarantees?''
+This allows them to use the logical relation to reason about arbitrary (untrusted) programs.
+We also want to be able to say something about arbitrary (untrusted) programs, but we also want to be able to say something about somewhat arbitrary trusted programs.
+In our setting, a trusted program is a well-formed, reasonable program that follows the \stktokens{} calling convention, and an untrusted program is an arbitrary well-formed program.
+In order for \stktokens{} to work, it needs access to return seals, but we cannot allow untrusted programs access to the return seals.
+A value relation based on what an adversary could do should permit return seals, so such a relation cannot be used to reason about trusted programs.
+For this reason, we define two value relations a trusted $\lrvtrusted$ and an untrusted $\lrv$.
+It is safe to give trusted programs anything that is safe for an untrusted program, so the trusted value relation is defined as a super set of the untrusted value relation.
+
+In Figure~\ref{fig:value-relation}, we have sketched the two value relations.
+This shows that for the most part, words on \srccm{} are related to words on \trgcm{} that are syntactical identical.
+The only exception is stack pointers on \srccm{} that are related to linear capabilities on \trgcm{}.
+Note that the return pointers of \srccm{} are not related to anything as it is never safe for any program, trusted or not, to have them.
+The \srccm{} return pointers should only occur under a seal, and they should only be used to jump to in which case the \srccm{} semantics transforms them to the capabilities they correspond to.
+
+The value relation is defined in terms of a number of auxiliary definitions.
+In Paragraph~\ref{par:perm-cond} and Paragraph~\ref{par:standard-regions} we present the auxiliary definitions.
+In Paragraph~\ref{par:untrusted-val-rel} and Paragraph~\ref{par:trusted-val-rel}, we present the untrusted and trusted value relation, respectively.
+
+% The old section:
+% Note first how we have two value relations, whose definitions are sketched in Figure~\ref{fig:value-relation}.
+% The difference is that the untrusted value relation $\lrv(W)$ does not just express that the two values are related, but also that they are safe to pass to an untrusted adversary, i.e.
 they cannot be used to break LSE and WBCF.
-The trusted value relation does not have the latter requirement and is a superset of the former.
-
-Both relations trivially include numbers $(i,i)$ which are always related to themselves.
-The untrusted value relation also includes stack pointers and the underlying linear capability (with the same (non-executable) permission, range of authority, and current address), as well as syntactically equal memory capabilities, seals and sealed values, all under certain conditions involving the world $W$ and the capability's properties.
-
-Roughly, for stack capabilities, the omitted condition requires that the world contains a $\spatialo$ region governing this part of the stack.
-For memory capabilities $((\perm,\lin),\baddr,\eaddr,\aaddr)$, a region in the world must govern memory $[\baddr,\eaddr]$, either $\spatialo$ or $\pure$, depending on the linearity $\lin$ of the capability.
-If the capability is executable ($\perm \in \{\rx,\rwx\}$), then we additionally require that the governing region is a code region and that the two capabilities are related $\mathrm{jmp}$ targets, as expressed by the relation $\lre(W)$, in any future world (see below).
-
-Seals allocated to trusted code are related to themselves only by $\lrvtrusted(W)$, but other seals are in both value relations.
-Sealed values are in both relations essentially when the sealed values satisfy the relation that was registered for the seal in a region of the world.
-Additionally, when they are combined with any other pair of values related by that relation, they must be related as $\txjmp{}{}$ targets (i.e. in $\lrexj(W)$).
-Finally, capabilities to code memory are related to themselves in the trusted value relation ($\lrvtrusted(W)$) when there is an appropriate code region in the world.
-They are not in the untrusted value relation because the code memory contains copies of the return seals used by the code, which must not end up in the hands of an adversary.
+% The trusted value relation does not have the latter requirement and is a superset of the former.
+
+% Both relations trivially include numbers $(i,i)$ which are always related to themselves.
+% The untrusted value relation also includes stack pointers and the underlying linear capability (with the same (non-executable) permission, range of authority, and current address), as well as syntactically equal memory capabilities, seals and sealed values, all under certain conditions involving the world $W$ and the capability's properties.
+
+% Roughly, for stack capabilities, the omitted condition requires that the world contains a $\spatialo$ region governing this part of the stack.
+% For memory capabilities $((\perm,\lin),\baddr,\eaddr,\aaddr)$, a region in the world must govern memory $[\baddr,\eaddr]$, either $\spatialo$ or $\pure$, depending on the linearity $\lin$ of the capability.
+% If the capability is executable ($\perm \in \{\rx,\rwx\}$), then we additionally require that the governing region is a code region and that the two capabilities are related $\mathrm{jmp}$ targets, as expressed by the relation $\lre(W)$, in any future world (see below).
+
+% Seals allocated to trusted code are related to themselves only by $\lrvtrusted(W)$, but other seals are in both value relations.
+% Sealed values are in both relations essentially when the sealed values satisfy the relation that was registered for the seal in a region of the world.
+% Additionally, when they are combined with any other pair of values related by that relation, they must be related as $\txjmp{}{}$ targets (i.e. in $\lrexj(W)$).
+% Finally, capabilities to code memory are related to themselves in the trusted value relation ($\lrvtrusted(W)$) when there is an appropriate code region in the world.
+% They are not in the untrusted value relation because the code memory contains copies of the return seals used by the code, which must not end up in the hands of an adversary.
 
@@
 \end{align*}
-\caption{Sketches of the trusted and untrusted value relation.}
+\caption{Sketches of the trusted and untrusted value relation.
+The untrusted and trusted value relation both relates \srccm{} and \trgcm{} words.
+The untrusted value relation $\lrv$ relates words that are safe to give to untrusted programs and $\lrvtrusted$ relates words that are safe to give to trusted programs.}
 \label{fig:value-relation}
@@
 \paragraph{Permission based conditions}
+\par{par:perm-cond}
 \[
@@
 \paragraph{The untrusted value relation}
+\label{par:untrusted-val-rel}
 \[
@@
 \paragraph{The trusted value relation}
+\label{par:trusted-val-rel}
 \[
@@
 \end{lemma}
-
-\subsubsection{Observation relation}
-\[
-  \lro[\preceq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
-    \begin{array}{l}
-      \forall i \leq n \ldotp \\
-      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
-    \end{array}
-\right\}
-\]
-\[
-  \lro[\succeq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
-    \begin{array}{l}
-      \forall i \leq n \ldotp \\ 
-      \quad (\ms_T ,\reg_T) \term[i] \\\qquad\Rightarrow \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm{\ta,\stkb}
-    \end{array}
-\right\}
-\]
 
@@
 \end{align*}
-
+\end{jversion}
 \subsection{Fundamental Theorem}

### 2019-05-21 fully-abstract-comp/ftlr-proof-cases.tex
[POPLjv] Added region paragraph

@@
 \end{itemize}
-from this, we get $S_{\var{read}} \subseteq \addressable{\linear,\pwheap{W}}$ and
+from this, we get $S_{\var{read}} \subseteq \addressable{\linear,\pwheap}$ and
 $R_{\var{read}} : S_{\var{read}} \fun \Worlds$ such that
@@
 
-By assumption $\pwheap(r)$ is address-stratified which means that for any $\hat{W}$ and any $\ms$ and $\ms'$ where $\npair[n-1]{(\ms,\ms')} \in \pwheap(r)$ we have $\dom(\ms)e= \dom(\ms') = \{a'\}$.
+By assumption $\pwheap(r)$ is address-stratified which means that for any $\hat{W}$ and any $\ms$ and $\ms'$ where $\npair[n-1]{(\ms,\ms')} \in \pwheap(r)$ we have $\dom(\ms) = \dom(\ms') = \{a'\}$.
 
@@
 the same domain, we cannot satisfy all the regions, so we must have
+% TODO something has gone wrong here. It is not clear to me from the history what happened.
 

### 2019-05-21 fully-abstract-comp/technical-report.tex
Add permissions to POPL19 technical report
Fix typos and omissions in POPL technical report
Fix small typo in POPL technical report
[POPLjv] Added region paragraph

@@
     w \in &\Word & \defeq & \ints \uplus \Caps\\
-    \permbnf \in& \Perm & \defbnf & \dots \\
+    \permbnf \in& \Perm & \defbnf & \noperm \mid \readonly \mid \rw \mid \rx
+                                    \mid \rwx \\
     &\linbnf & \defbnf & \linear \mid \normal \\
@@
 For the linearity encoding, we make similar assumptions: $\encLin{} : \Linear \fun \ints$ encodes linearity.
-$\encLin{}$ is injective and for all $\lin \in \Linear$ $\encPerm{\lin} \neq -1$ (as this is used as an error value).
+$\encLin{}$ is injective and for all $\lin \in \Linear$ $\encLin{\lin} \neq -1$ (as this is used as an error value).
 As a capabilities linearity stays the same, we do not need a decoding function for linearity.
@@
   (\spatial,H) \oplus (\spatial,H) =  & \; (\spatial,H) \\
+  \revoked \oplus \revoked = & \; \revoked \\
   (\spatialo,H) \oplus (\spatial,H) = & \; (\spatial,H) \oplus (\spatialo,H)\\
@@
   \inferrule{ }{(\spatial,H) \future (\spatial,H)}
+  \and
+  \inferrule{ }{(\pure,H) \future (\pure, H)}
 % \and
@@
 
-% \dominique{TODO: better name than ``reasonable'' for all of this.}
 Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
@@
 
-    By definition of $\lrv$, it suffices to show that $\npair{[b,e]} \in \readCond{\normal,W}$ and $\npair{[b,e]} \in \execCond{\normal,W}$.
+    By definition of $\lrv$, it suffices to show that $\npair{[b,e]} \in \readCond{\normal,W}$ and $\npair{[b,e]} \in \execCond{W}$.
     The former follows directly from Lemma~\ref{lem:xReadCond-outside-ta-implies-readCond}.
@@
   Follows easily from the definitions.
-  In the case for $\execCond{}$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
-  It suffices construct new $W_a'$ by taking $W_a$ but adding $\oplus_{a\in(A\setminus A')}$ to an $W_a(a)$ for an arbitrary $a \in A'$ to make this work.
+  % Note: the case for execCond where \lin = \linear does not occur
+  % In the case for $\execCond{}$ where $\lin = \linear$, we have a partition of the world into $W = \oplus_{a \in A} W_a(a)$ that we need to convert into a partition $W = \oplus_{a \in A'} W_a'(a)$.
+  % It suffices construct new $W_a'$ by taking $W_a$ but adding $\oplus_{a\in(A\setminus A')}$ to an $W_a(a)$ for an arbitrary $a \in A'$ to make this work.
 \end{proof}

### 2019-05-21 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Added region paragraph

@@
 \]
+For readability we will sometimes use $\spao$ as short for $\spatialo$, $\spa$ as short for $\spatial$, and $\pur$ as short for $\pure$.
 
@@
 
+In this section, we will from time to time refer to safety of a capability or a word.
+In some sense, our logical relation actually ends up as the definition of safety, so when we refer to a capability as \emph{safe} it is in an informal sense where it means that the capability cannot be used break memory invariants.
+
 In Figure~\ref{fig:value-relation}, we have sketched the two value relations.
@@
 The value relation is defined in terms of a number of auxiliary definitions.
-In Paragraph~\ref{par:perm-cond} and Paragraph~\ref{par:standard-regions} we present the auxiliary definitions.
-In Paragraph~\ref{par:untrusted-val-rel} and Paragraph~\ref{par:trusted-val-rel}, we present the untrusted and trusted value relation, respectively.
+In the following, we introduce a number of \emph{standard regions} that express common requirements on memory.
+Based on the standard regions, we define what we call \emph{permission based conditions}, conditions that a capability with a specific permission must satisfy to be safe.
 
@@
 
-\paragraph{Permission based conditions}
-\par{par:perm-cond}
-\[
-  \addressable{\lin,W} =
-  \begin{cases}
-    \{ r \mid W(r) = (\pure,\_) \} & \text{if $\lin = \normal$} \\
-    \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
-  \end{cases}
-\]
-
-\[
-  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
-    \begin{array}{l}
-      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
-      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
-      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
-    \end{array}
-  \right\}
-\]
-\[
-  \stackReadCond{W} = \left\{ \npair{A} \middle| 
-    \begin{array}{l}
-      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
-      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
-      \qquad \forall r \in S \ldotp |R(r)| = 1\\
-      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
-    \end{array}
-  \right\}
-\]
-where $\iota_A$ is a standard region defined in Section~\ref{par:standard-regions}.
-\[
-  \xReadCond{W} = \left\{ \npair{A} \middle| 
-    \begin{array}{l}
-      \exists r \in \addressable{\normal, \pwheap} \ldotp \\
-      \qquad \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\wedge\\
-      \qquad \dom(\code) \supseteq A 
-    \end{array}
-  \right\}
-\]
-
-\begin{definition}
-  \label{def:address-stratified}
-  We say that a region $\iota = (\_,H,\_)$ is address stratified iff
-  \[
-    \begin{array}{l}
-      \forall n, \src{\ms_S},\ms_T,\src{\ms_S'},\ms_T',s,\hat{W}\ldotp \\
-      \quad \npair{\stpair{\ms}{\ms}}, \npair{\stpair[.]{\ms_S'}{\ms_T'}} \in H \; \hat{W} \wedge \\
-      \quad \dom(\src{\ms_S}) = \dom(\ms_T) = \dom(\src{\ms_S'}) = \dom(\ms_T') \\
-      \quad \Rightarrow \\
-      \qquad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\src{\ms_S}\update{\aaddr}{\src{\ms_S'}(\aaddr)},\ms_T\update{\aaddr}{\ms_T'(\aaddr)})} \in H \; \hat{W}
-    \end{array}
-  \]
-\end{definition}
-\[
-  \writeCond{\lin,W} = \left\{ \npair{A} \middle| 
-    \begin{array}{l}
-      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
-      \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
-      \qquad (\lin = \linear \Rightarrow \forall r\ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
-      \qquad \quad \pwheap(r) \text{ is address-stratified}
-    \end{array}
-  \right\}
-\]
-where $\iota_A$ is a standard region defined in Paragraph~\ref{par:standard-regions}.
-\[
-  \stackWriteCond{W} = \left\{ \npair{A}) \middle| 
-    \begin{array}{l}
-      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
-      \quad \exists R : S \fun \powerset{\nats} \\
-      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
-      \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
-      \qquad \quad \pwfree(r) \text{ is address-stratified}
-    \end{array}
-  \right\}
-\]
-where $\iota_A$ is a standard region defined in Paragraph~\ref{par:standard-regions}.
-\begin{align*}
-  \execCond{W} &=
-  \left\{ \npair{A} \middle|
-    \begin{array}{l}
-      \forall n' < n, W' \future \purePart{W}, \aaddr \in [\baddr',\eaddr'] \subseteq A \ldotp\\
-      \quad \npair[n']{\stpair[.]{((\rx,\normal),\baddr',\eaddr',\aaddr)}{((\rx,\normal),\baddr',\eaddr',\aaddr)}} \in \lre(W')
-    \end{array}
-    \right\}
-\end{align*}
-
 \paragraph{Standard regions}
 \label{par:standard-regions}
-Standard region:
-\[
-  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc) , v \in \{\spa,\spao\}
-\]
-for readability, we use $\spao$ short for $\spatialo$, $\spa$ short for $\spatial$, and $\pur$ as short for $\pure$.
-\[
-  \stdreg{A,\gc}{\pur} \defeq (\pur,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
-\]
-where $H^{\mathrm{std},\square}_A$ is defined as follows:
+The notion of regions we defined in Section~\ref{subsec:worlds} is general enough to allow a wide variety of regions.
+There are, however, some regions that may seem more natural or standard than others.
+In particular, when it comes to capability safety, it seems natural to have a region that requires everything in memory to be safe.
+This is exactly what we refer to as a \emph{standard region} because we usually define a region like that along with a logical relation.
+
+We define a $\pure$, $\spatial$, and $\spatialo$ standard. They all have the same invariant which is defined as follows:
 \[
@@
       \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
-      \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\
+      \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\ % TODO there should be a later on the right side of the equality?
       \quad \forall \aaddr \in A \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
@@
 \]
-
-\[
-  \stareg[\stpair{\ms}{\ms},\gc]{v,\square} = (v,H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc)
-\]
-\[
-  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc = \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
-    \begin{array}{l}
-      \exists S : \dom(\ms) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in \dom(\ms) \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+The standard region invariant requires the memory segment pairs to have a specific address space $A$.
+Further, the two memory segments must contain words from the untrusted value relation.
+The memory segments may contain linear capabilities, so we must distribute the ownership of the world between each memory cell which the function $S$ takes care of.
+Note that the invariant takes a $\hat{W}$ from $\Wor$ as argument which means that we must apply the isomorphism $\xi$ before the world can be used.
+Using this invariant, we define the standard $\spatial$ and $\spatialo$ regions as follows:
+\[
+  \stdreg{A,\gc}{v} \defeq (v,H_A^{\mathrm{std},\square} \; \gc) , v \in \{\spatial,\spatialo\}
+\]
+and the standard $\pure$ regions as follows
+\[
+  \stdreg{A,\gc}{\pure} \defeq (\pur,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
+\]
+Note that the standard $\pure$ region puts no requirements on seals.
+
+Sometimes we need to know that the contents of a memory segment stays the same.
+For instance, the contents of encapsulated stack frames don't change which we need to be able to rely on.
+To express this, we define a \emph{static region}.
+The static region is parameterised with a memory segment pair which is the only memory segment pair the region accepts.
+The memory invariant is defined as follows
+\[
+  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \defeq \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+    \begin{array}{l}
+      \dom(\ms_S) = \dom(\ms_T) \wedge \\
+      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
+      \quad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
 \]
-
-\[
-  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure,
-H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc,
-H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
-\]
-
+The region also requires the static memory to contain words from the untrusted value relation.
+Which just means that the stack shouldn't be used for return seals, closure seals, and code pointers for trusted code.
+With the memory invariant, we define the static region as follows:
+\[
+  \stareg[\stpair{\ms}{\ms},\gc]{v,\square} \defeq (v,H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc) , v \in \{\spatial,\spatialo\}
+\]
+A $\pure$ static region can be defined in a similar fashion to that of the standard region.
+
+In our results, we assume that we execute well-formed components which allows us to make certain assumptions about the code we execute.
+We express these assumption in what we call a \emph{code region}.
+The invariant of the code region is defined as
 \begin{multline*}
@@
 \end{multline*}
-
+The code region is a lot more restrictive that the standard region.
+It basically only allows one memory segment, namely $\code$ padded with zeroes to make sure that two capabilities cannot be spliced to cause unintended control-flow.
+We will use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
+The assumptions we can make on the code depends on whether it is part of a trusted or untrusted component.
+This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
+This is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then it is trusted.
+On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then it is untrusted.
+In either case, the words should be in the value relation with respect to the $\purePart{}$ of the world which means that the code memory cannot contain linear capabilities.
+% TODO promote $\purePart{}$ from appendix to paper - used a few times.
+
+\stktokens{} rely on proper seal usage to guarantee well-bracketed control-flow and local state encapsulation.
+This means that components must use return and closure seals for their intended purpose.
+The code region has a seal invariant $H^\mathrm{code,\square}_\sigma$ to guarantee that the return and closure seals of the region are used correctly.
+The seal invariant is displayed in Figure~\ref{fig:code-reg-seal-inv}.
+The return seals $\sigrets$ in a code region should only be used to seal return pointers.
+That is on \srccm{} they should only be used to seal $\retptrc{}$ and $\retptrd{}$.
+\begin{figure}
+  \centering
   \begin{multline*}
   H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
-  (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  (\ta,\stkb,\_,\gsigrets) \; \sigma \; \hat{W} \defeq \\
   \begin{array}[t]{l}
@@
     \end{array}
-      \right\} \uplus \\
+      \right\} \cup \\
 \left\{
@@
         \dom(\code) \subseteq \ta \tand\\
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \\
+        \quad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
@@
   \end{array}\\
-    \text{ if } \sigma \in \sigrets\\
+    \text{ for } \sigma \in \sigrets
 \end{multline*}
-and
 \begin{multline*}
   H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
-  (\ta,\stkb) \; \sigma \; \hat{W}= \\
+  (\ta,\stkb,\gsigcloss,\gsigrets) \; \sigma \; \hat{W} \defeq \\
   \left\{
@@
   \right\}\\
-  \text{ if } \sigma \in \sigcloss
+  \text{ for } \sigma \in \sigcloss
 \end{multline*}
+
+\caption{The seal invariant for code regions.}
+\label{fig:code-reg-seal-inv}
+\end{figure}
+% Code return pointer
+If we allowed any $\retptrc$ to be sealed, then we could not be sure that it comes from a call even though it should only be possible to get a return pointer from a call.
+For this reason, we require that the \srccm{} return pointer actually points to the first address after a call.
+For a \trgcm{} capability related to a \srccm{} code return pointer, we do not require that it points to the first address after the call because it needs to execute the return instructions first.
+
+% Data return pointer
+For sealed data return pointers, we need to know that the world contains a region that governs the local stack frame.
+That is, there should be a static region with the contents of the stack frame.
+The fact that it is static means that the contents will remain the same.
+The region that governs the stack frame must come from the private world which means that it is paired with a return address.
+We require that the return address actually corresponds to a return address for a call in the code.
+
+% Closure seals
+Unlike return seals, both trusted and untrusted components can have closure seals.
+For untrusted components (components with their code address space disjoint from the trusted address space), we allow everything in the untrusted value relation to be sealed.
+Intuitively, untrusted components are assumed to have access to words from the untrusted value relation, but we cannot know how the words are used, so we need to assume that they may be sealed.
+% TODO trusted components
+Trusted components only use closure seals for sealed capability pairs that represent actual closures.
+The code capability for a closure must point to the code memory because it is the only part of memory that is executable.
+It is not safe for untrusted components to have a capability for a trusted components code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
+While it is not safe to give a bare capability for a trusted components code memory, it can be perfectly safe to give a sealed capability for a trusted components code.
+For this reason, the seal invariant allows executable capabilities from the trusted value relation to be sealed with a closure seal.
+When it comes to the data capability of a closure, we just require that it comes from the untrusted value relation because the trusted value relation contains nothing that makes sense to seal as the data capability (we return to the specific contents of the two value relations in a later paragraph).
+
+With the memory invariant and seal invariant in hand, we define the code region as follows:
+\[
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure, H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc, H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+\]
+The code region is $\pure$ because it needs to contain a seal invatiant and because we assume that code pointers are non-linear.
+
+\paragraph{Permission based conditions}
+\label{par:perm-cond}
+% TODO add brief intro to paragraph
+
+The world can be seen as an authority specification which means that it dictates what kind of capabilities can address a certain part of memory.
+Specifically, linear capabilities can only address memory governed by a $\spatialo$ region, and non-linear capabilities can only address memory governed by a $\pure$ region.
+All the permission based condition we define project the regions that the capability may address from the world.
+The following $\addressable$ function takes care of the projection:
+\[
+  \addressable{\lin,W} \defeq
+  \begin{cases}
+    \{ r \mid W(r) = (\pure,\_) \} & \text{if $\lin = \normal$} \\
+    \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
+  \end{cases}
+\]
+
+% TODO introduce read condition
+We capture the essence of what it means for a capability with read permission to be safe in the condition $\readCond{}$.
+% each of the regions upper bounded by std region
+The main purpose of $\readCond{}$ is to make sure that only safe words can be read from the memory the read capability governs.
+This is done by putting an upper bound on what requirements a region can impose on the memory segments governed by the capability.
+In particular a region that governs the memory a read capability has access to can at most allow safe values to be read.
+Without this requirement, a read capability could potentially be used to break memory invariants if were used to read capabilities that has the authority to break memory invariants.
+The condition is defined as follows
+\[
+  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad (\lin = \linear \Rightarrow \forall r \in S \ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
+    \end{array}
+  \right\}
+\]
+% splicing and what part of the range each region governs
+The $\readCond{}$ is compatible with all the operations that can be performed on capabilities.
+This means that if two capabilities, for which $\readCond{}$ holds, are spliced together, we should be able to establish that $\readCond{}$ holds for the resulting capability.
+To support this, we require the presence of a set of regions $S$ that governs the addresses the capability has authority over rather than just a single region.
+If we need to establish the $\readCond{}$ after a splice, we can simply use union of the regions that witnessed the $\readCond{}$ of the two individual capabilities.
+We also need to support splitting which is no problem for non-linear capabilities as the same $\pure$ region can be used to establish the $\readCond{}$ for multiple non-linear capabilities.
+On the other hand, a $\spatialo$ region can only be used to establish the $\readCond{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
+None the less, we need to support arbitrary splitting of linear capabilities, which means that $\readCond{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond{}$.
+This is why, $\readCond{}$ requires all regions to only govern one address when the capability is linear.
+This means that after a split, the authority of the regions for the bottom half of the split result can go to one capability and the rest can go to the top half.
+
+The upper bound on requirements that can be imposed on the memory the read capability has access to is specified on all of the regions that govern the memory.
+The upper-bound is specified in terms of the standard region $\stdreg{R(r),\gc}{\pur}$.
+That is, the memory segments permitted by the region should be an $n$-subset (a subset that only takes into account elements with an index up to, but not including, $n$).
+This means that everything the region permits must at the very least be safe.
+
+% TODO is there a point to be made about limited write for reading of linear capabilities?
+% When a linear capability is read from memory, the source address must be cleared.
+% This means that a linear capability with read permission also has a very limited form of write authority which only permits it to write zeroes.
+
+Similarly to $\readCond{}$, we define a condition that captures the essence of what it means for a capability with writer permission to be safe.
+We call this condition $\writeCond{}$.
+A capability with write permission can be used to write to memory.
+The question is, what can we safely allow to be written to memory without any memory invariants being broken.
+It may come as a surprise that the answer to this is anything - even words that are unsafe.
+The intuition behind this is that if you managed to write something that can break memory invariants, it would not be possible to read it back again as you need read permission for that, and if the capability had read permission, then $\readCond{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants, after all if one was obtained, memory invariants could be broken.
+However, the $\writeCond{}$ tries to capture the essence of safety and in principle it is safe to write an unsafe capability that cannot be read back.}.
+We do, however, impose a bound on what should be permitted to be written and that is a lower bound because it should at least be allowed to write anything that is safe.
+
+Before we get to the definition of $\writeCond{}$, we need to introduce one auxiliary definition.
+When a write capability makes a change to memory, the change must be permitted by the regions that govern the memory it has authority over for it to be safe.
+Given a write capability it should be possible to transform the memory from one allowed by the region to another also allowed by the region.
+However, only one memory address can be updated at a time, so in order to make such a transformation one has to update one address at a time.
+This should not break any memory invariants, so we require all the memories that we could go through to make the transformation to also be permitted by the region.
+This is captured by address stratification.
+\begin{definition}
+  \label{def:address-stratified}
+  We say that a region $\iota = (\_,H,\_)$ is address stratified iff
+  \[
+    \begin{array}{l}
+      \forall n, \src{\ms_S},\ms_T,\src{\ms_S'},\ms_T',s,\hat{W}\ldotp \\
+      \quad \npair{\stpair{\ms}{\ms}}, \npair{\stpair[.]{\ms_S'}{\ms_T'}} \in H \; \hat{W} \wedge \\
+      \quad \dom(\src{\ms_S}) = \dom(\ms_T) = \dom(\src{\ms_S'}) = \dom(\ms_T') \\
+      \quad \Rightarrow \\
+      \qquad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\src{\ms_S}\update{\aaddr}{\src{\ms_S'}(\aaddr)},\ms_T\update{\aaddr}{\ms_T'(\aaddr)})} \in H \; \hat{W}
+    \end{array}
+  \]
+\end{definition}
+We define $\writeCond{}$ as follows
+\[
+  \writeCond{\lin,W} \defeq \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats}\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
+      \qquad (\lin = \linear \Rightarrow \forall r \in S \ldotp |R(r)|  = 1) \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
+      \qquad \quad \pwheap(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\]
+The definition of $\writeCond{}$ is very similar to $\readCond{}$.
+Support for split and splice is done in the same way, and the bound is defined in terms of the standard region.
+
+The $\readCond{}$ and $\writeCond{}$ specifically uses the heap sub-world which means that it can only be used for heap capabilities.
+This means that we cannot use it for stack capabilities.
+To take care of stack capabilities, we define two more conditions a $\stackReadCond{}$ and $\stackWriteCond{}$.
+The two new condition are essentially the same as the $\readCond{}$ and $\writeCond{}$ except that they use the free stack sub-world and assume that the capability is linear as all stack capabilities are linear.
+Note that we do not have any condition that talks about the stack-frames sub world because we should never have a capability that allows us to directly read or write from that part of memory.
+\[
+  \stackReadCond{W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
+      \qquad \forall r \in S \ldotp |R(r)| = 1\\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
+    \end{array}
+  \right\}
+\]
+\[
+  \stackWriteCond{W} = \left\{ \npair{A}) \middle| 
+    \begin{array}{l}
+      \exists S \subseteq \addressable{\linear, \pwfree} \ldotp \\
+      \quad \exists R : S \fun \powerset{\nats} \\
+      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge \\
+      \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
+      \qquad \quad \pwfree(r) \text{ is address-stratified}
+    \end{array}
+  \right\}
+\]
+The final permission, we define conditions for is the execute permission.
+We define two conditions $\execCond{}$ and $\xReadCond{}$.
+The $\execCond{}$ captures what operations an execute-capability can be used for, i.e.\ execution.
+The $\xReadCond{}$ captures some additional read assumptions we can make on a capability when we know the capability is executable.
+
+The $\execCond{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
+We later define the $\lre$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
+An executable capability can have its range of authority shrunk or its current address changed which changes what it does and thus potentially whether it keeps memory invariants.
+For this reason, the condition requires that any capability with a derived range of authority and a current address that range is safe to use for execution.
+The $\execCond{}$ is quantified over all future worlds of the $\purePart{}$ of $W$. The $\purePart{}$ function strips the world of all spatial ownership.
+This means that the capability cannot depend on linear capabilities, and when we define the logical relation, we even require the executable capability to not be linear.
+It is not clear whether a linear executable capability would even be useful because there is no way to move it out of the $\pcreg$-register without crashing the execution.
+This may sound like an ideal option for constructing something that can be executed once, however, most programs rely on loading other capabilities or seal sets for the program counter which is not possible.
+\begin{align*}
+  \execCond{W} &=
+  \left\{ \npair{A} \middle|
+    \begin{array}{l}
+      \forall n' < n, W' \future \purePart{W}\ldotp \forall b',e'\ldotp \forall \aaddr \in [b',e'] \subseteq A \ldotp\\
+      \quad \npair[n']{\stpair[.]{((\rx,\normal),\baddr',\eaddr',\aaddr)}{((\rx,\normal),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+    \end{array}
+    \right\}
+\end{align*}
+
+% TODO Motivate the read execute condition (do this when description of code region done)
+\[
+  \xReadCond{W} = \left\{ \npair{A} \middle| 
+    \begin{array}{l}
+      \exists r \in \addressable{\normal, \pwheap}, \code \ldotp \\
+      \qquad \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\wedge\\
+      \qquad \dom(\code) \supseteq A 
+    \end{array}
+  \right\}
+\]
+% TODO Stack not executable, so no stack readx or stack exec cond.
 

### 2019-05-21 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add description of readXCond

@@
 Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
-The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.
+The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.
 
@@
 \end{align*}
-
-% TODO Motivate the read execute condition (do this when description of code region done)
+The $\readCond{}$ condition by itself allows many different regions and thus potentially many different memories.
+However, when we have a read capability with execute permission, we know that the capability must point to a piece of this code memory.
+For this reason, we define the $\xReadCond{}$ to capture the additional assumptions that we can make when a capability is executable.
 \[
@@
 \]
-% TODO Stack not executable, so no stack readx or stack exec cond.
-
+The $\xReadCond{}$ requires that the memory segment an executable capability has authority over is governed by a code region.
+Note that we do not define $\execCond{}$ and $\xReadCond{}$ for the stack because the stack is not executable.
 \paragraph{The untrusted value relation}

### 2019-05-21 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Update todos

@@
 Intuitively, untrusted components are assumed to have access to words from the untrusted value relation, but we cannot know how the words are used, so we need to assume that they may be sealed.
-% TODO trusted components
 Trusted components only use closure seals for sealed capability pairs that represent actual closures.
@@
 \label{par:untrusted-val-rel}
+% TODO write section
 \[
@@
 \label{par:trusted-val-rel}
+% TODO write section
 \[
@@
 \subsubsection{Expression relations}
+% TODO write section
 % Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
@@
 \subsection{Fundamental Theorem}
+% TODO rewrite section
+% Include the actual FTLR
 An important lemma in our proof of full abstraction of the embedding of \srccm{} into \trgcm{}, is the fundamental theorem of logical relations (FTLR).
@@
 After that first step, the induction hypothesis is used for the rest of the execution.
+
+\subsection{Related Components}
+% TODO write section
+\begin{align*}
+  \lrcomp(W) &=
+  \left\{\begin{aligned}
+     &\npair{\var{comp},\var{comp}} \;\mid \;\\
+      &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+      &\qquad\text{For all } W' \future W \ldotp \\
+      &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+      &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
+      &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+  \end{aligned}
+    \right\}\\
+  &\cup \left\{
+    \begin{multlined}
+     \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
+     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
+     \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
+       \end{multlined}
+    \right\} 
+\end{align*}
+% TODO include central lemmas?
+
+\subsection{Related Execution Configuration}
+% TODO write section
+\[
+  \lrec(W) = \left\{ \npair{\left(
+        \arraycolsep=0pt\array{l}
+        (\ms_S,\reg_S,\stk,\ms_\stk),\\
+        (\ms_T,\reg_T)\endarray\right)} \middle|
+    \begin{array}{l}
+      \gc = (\ta,\stkb) \tand\\
+      \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
+      \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
+      \reg_S(\pcreg) \neq \retptrc(\_) \wedge 
+      \reg_S(\rdata) \neq \retptrd(\_) \wedge \\
+      \nonExec{\reg_S(\rdata)} \wedge
+      \nonExec{\reg_T(\rdata)}\\
+      \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
+      \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)
+    \end{array}
+            \right\}
+\]
+
+% TODO Include central lemmas?
 

### 2019-06-16 fully-abstract-comp/technical-report.tex
[POPLjv] More work on the POPL journal version

@@
                        \begin{cases}
-                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (pure,\var{sm})\\
+                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (\pure,\var{sm})\\
                          (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
@@
                        \begin{cases}
-                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (pure,\var{sm})\\
+                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (\pure,\var{sm})\\
                          (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\

### 2019-06-16 popl-2019-journal-v/appendix.tex
[POPLjv] Add world erasure to appendix
[POPLjv] More work on the POPL journal version

--- 2019-04-05 popl-2019-journal-v/appendix.tex
@@
 
+  \section{World definitions}
+  \begin{definition}
+    Given $W_{\mathrm{free}} \in \Worldfs$
+    \[
+\erase{W_{\mathrm{free}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{W_{\mathrm{free}}(r)} \in S\\
+      \bot
+    \end{array}
+  \right.
+\]
+
+  \end{definition}
+
+  \begin{definition}[Private stack sub-world erasure]
+    Given $W_{\mathrm{priv}} \in \Worlds$
+    \[
+\erase{W_{\mathrm{priv}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{\pregion{W_{\mathrm{priv}}(r)}} \in S \\
+      \bot
+    \end{array}
+  \right.
+\]
+  \end{definition}
+  \begin{definition}[Heap sub-world erasure]
+    Given $W_{\mathrm{heap}} \in \Worldh$
+    \[
+      \erase{W_{\mathrm{heap}}}{S} = \lambda r \ldotp \left\{
+        \begin{array}{l}
+          \prv{W_{\mathrm{heap}}(r)} \in S\\
+          \bot
+        \end{array}
+      \right.
+    \]
+  \end{definition}
+  \begin{definition}[World erasure]
+    Given world $(W_{\mathrm{heap}},W_{\mathrm{priv}},W_{\mathrm{free}})$ define world erasure as
+\[
+\erase{(W_{\mathrm{heap}},W_{\mathrm{priv}},W_{\mathrm{free}})}{S} =
+  \left( \erase{W_{\mathrm{heap}}}{S},\erase{W_{\mathrm{priv}}}{S},\erase{W_{\mathrm{free}}}{S} \right)
+\]
+\[
+\erase{W_{\mathrm{heap}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{W_{\mathrm{heap}}(r)} \in S\\
+      \bot
+    \end{array}
+  \right.
+\]
+\[
+\erase{W_{\mathrm{priv}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{\pregion{W_{\mathrm{priv}}(r)}} \in S \\
+      \bot
+    \end{array}
+  \right.
+\]
+\[
+\erase{W_{\mathrm{free}}}{S} = \lambda r \ldotp \left\{
+    \begin{array}{l}
+      \prv{W_{\mathrm{free}}(r)} \in S\\
+      \bot
+    \end{array}
+  \right.
+\]
+
+The $\activeReg{}$ function takes a world and filters away all the revoked regions, so
+\[
+  \activeReg{W} = \erase{W}{\spatial,\spatialo,\pure}
+\]
+  \end{definition}
+
 %%% Local Variables:

### 2019-06-16 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] More work on the POPL journal version

@@
 For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} are modelled by the worlds.
-In order to related \trgcm{} and \srccm{}, we need to model all the features of \srccm{} in the worlds which means we have to model:
+In order to related \trgcm{} and \srccm{}, we model all the features of \srccm{} in the worlds which means we have to model:
 \begin{itemize}
@@
 \end{itemize}
+The three kinds of memory are modelled by having three sub-worlds where each sub-world is its own little world in a traditional sense.
+Linearity is modelled by adding ownership to certain parts of the world, so capabilities can take ownership of part of the memory thus ensuring that they are the sole reference.
+The call stack is modelled by the memory satisfaction (the relation that decides whether two memories are related in the world) by ensuring that the memory is actually shaped like a stack
+Finally, the seals are modelled by seal invariants that make sure that seals only are used on permitted sealables.
+In the following, we present the world and go into details about how each of the four points are modelled.
 
@@
 We call the invariants regions because, as we will see later, they turn out to not be invariant.
-A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments ($\Rel{\MemSeg \times \MemSeg}$).
+A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
 Intuitively, we want to be able to say that two memory segments are related when their contents are related.
-That is, for every address in the two memory segments, the words that resides there must be related.
-In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now consider the intuition.
-Say we have two integers, then they are related if they are equal.
-If we instead have two capabilities, then they should also be related if they in some sense are equal, but what does it mean for capabilities to be ``equal''?
-Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably do the same computation.
+That is, for every address in the two memory segments, the words that reside there must be related.
+In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some the intuition.
+Say we have two integers, they are related if they are equal.
+If we instead have two capabilities, then they should also be related if they, in some sense, are equal.
+But what does it mean for capabilities to be ``equal''?
+Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably perform the same computation.
 As a capability points to a piece of memory, the authority of the capability depends on the contents of that memory.
-To say whether two capabilities are related, we must know the possible contents of the memory.
-This is exactly what our world express, so our regions will be world indexed, i.e.
+In other words to say whether two capabilities are related, we must know the possible contents of the memory.
+The world expresses the possible memory contents, so our regions are world indexed, i.e.
 \[
@@
 \end{gather*}
-
-If we just wanted to model the memory, then the above regions would suffice, but we also need the world to model linearity and seals.
-In the following subsections, we will gradually introduce the necessary components for our worlds.
-
-\subsubsection{Linearity and world joins}
+\subsubsection{Linearity}
 % Linearity
-The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they can reference.
-In order to model this uniqueness, we need to keep track of which parts of memory that are uniquely referenced and make sure that only one capability references the unique parts.
+The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they can reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
+In order to model this uniqueness, we need to keep track of which parts of memory that are uniquely referenced and make sure that only one linear capability references the unique parts.
 %% Bookkeeping done by worlds
@@
 If a memory segment is governed by a shared region, then non-linear capabilities may reference it.
-On the other hand, if a memory segment is governed by a spatial region, then only linear capability may reference it.
+On the other hand, if a memory segment is governed by a spatial region, then only a linear capability may reference it.
 %%% spatial/spatial_owned
@@
   \begin{array}{l}
-    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
-    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})\uplus \\ 
+    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
   \end{array} \right.
@@
 \]
+% TODO is the below notation even used?
 For readability we will sometimes use $\spao$ as short for $\spatialo$, $\spa$ as short for $\spatial$, and $\pur$ as short for $\pure$.
@@
 However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
-Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together constitute the sealed return pair.
+Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
 In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
@@
 
+Given a world, we sometimes want to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
+This is expressed by stripping the world of all its ownership, so the only active regions are the $\pure$ regions.
+We refer to this as the pure part of the world and define the function $\purePart{}$ which removes all ownership.
+\begin{definition}[The pure part of a world]
+  \label{def:purePart}
+  For any world $W$, we define
+  \begin{align*}
+    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
+    \purePart{W_\var{heap}} &\defeq
+                       \begin{cases}
+                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (\pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_\var{heap}(r) = \revoked\\
+                       \end{cases}\\
+    \purePart{W_\var{priv}} &\defeq
+                       \begin{cases}
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\opc)\\
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\opc)\\
+                         (\revoked,\opc) & \text{if } W_\var{priv}(r) = (\revoked,\opc)\\
+                       \end{cases}\\
+    \purePart{W_\var{free}} &\defeq
+                       \begin{cases}
+                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (\pure,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\
+                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
+                         \revoked & \text{if } W_{\var{free}}(r) = \revoked\\
+                       \end{cases}
+  \end{align*}
+\end{definition}
 \subsubsection{Seals}
@@
 \end{multline*}
+
+We will also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it i unambiguous.
 
@@
 The $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
-% Spatial -> spatialo ~ affine ?
+% TODO describe why "Spatial -> spatialo" ~ affine or at least why we have it?
 \todo[inline]{Lau: explain why $\spatialo \future \spatial$ gives us affine capabilitites.
@@
 \end{mathpar}
-The relation says that each of the three worlds must be an extension of the past world and for each of the existing regions that must be a future region.
+The relation says that each of the three worlds must be an extension of the past world and each of the existing region must have a future region.
 % injective function
@@
 
-% TODO monotonicity of regions
+The memory invariants are world dependent, but the worlds may change according to the future world relation, so what does that mean for the memory invariants?
+The memory invariants describe what certain parts of memory looks like.
+In future worlds, part of the memory may change but for the memory invariant to be invariant, it should keep relating the same memory segments even when the rest of the world changes.
+For this reason, we require the world-indexed memory invariants to be monotone in the world.
+This means that a pair of memory segments that are related now will stay related in future worlds.
+Our spatial regions now look like this:
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\revoked \}
+  \end{array} \right.
+\]
+We make a similar change for $\Regionh$.
+% TODO update terminology: memory invariant, H, and seal invariant, H_\sigma
 
@@
 \label{subsubsec:rec-dom-eq}
+% TODO consider new structure (the below)
+% First introduce c.o.f.e.s and the structures you can freely use to get a solution for your recursive domain equation
+% Then introduce all the specialized gunk necessary for our worlds
+%
 %solving the recursive domain equation means that we can present the actual definitions
@@
 % Move to a different domain
-Therefore, we need to move to a different domain with enough structure for a solution to exist for such an equation.
+Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
 We move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
 Intuitively, a c.o.f.e. is a set with some extra structure.
-Specifically, c.o.f.e.'s have a step-index which provide sufficient structure for us to solve recursive domain equations.
+Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
+which provide sufficient structure for us to solve recursive domain equations.
 In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.
@@
 \noindent One can think of the indexes on the family of equivalences as a measure of precision.
-The larger $n$ is, the more refined the equivalence may be.
-On the other hand as $n$ decreases, the equivalence becomes more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
+The larger $n$ is, the more refined the equivalences are.
+On the other hand as $n$ decreases, the equivalences become more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
 At index 0, all precision is lost and the equivalence cannot distinguish anything.
@@
   Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
-  elements of $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
+  elements in $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
   \begin{align*}
@@
 \noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
+% TODO add comment about standard constructions (x, +, fun, later, etc) and fixed-points (what are the conditions?
+Contractive, and?).
+Mention that it does not suffice to define our worlds, so we need to have some specialized worlds.
 However, this is not the only structure we would like to impose on our sets.
@@
 \begin{definition}[Uniform relation]
-  Given a relation $R \subseteq X \times Y$, we define $\URel{R} \subset \powerset{\nats \times R}$ as
+  Given a relation $R \subseteq X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
   \[
@@
   \[
-    \erase{A}{n} \defeq \{(k,a) \in A \mid k < n\}
+    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
   \]
@@
   \[
-    A \nequal B \text{ iff } \erase{A}{n} = \erase{B}{n}
+    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
   \]
 \end{definition}
-The $n$-equivalences of the uniform relation basically forgets everything from the step and up and requires the rest to be equal.
+The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step and up and requires the rest to be equal.
 This means that everything is equal at step $0$ as everything is left out.
 
-Next, we define the c.o.f.e.\ of monotone and non-expansive functions.
-%% Monotone, non-expansive
-\begin{lemma}[The c.o.f.e. of monotone and non-expansive functions]
-  Given c.o.f.e.'s $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future_W\right)$ and $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ and a preorder $\future$ on $W$, define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ as follows
-  \[
-    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
-  \]
-\end{lemma}
-Next, we define the product c.o.f.e.
-%% Product
-\begin{lemma}[Product c.o.f.e.]
-  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \right)$
-  where the family of equivalences are defined as follows
-  \[
-    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
-  \]
-\end{lemma}
-Similarly to the product c.o.f.e. we have a sum c.o.f.e.
-%% Sum
-\begin{lemma}[Sum c.o.f.e.]
-  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \right)$ define the sum c.o.f.e.\ as $\left(X \cup Y, \left(\nequal\right)_{n=0}^{\infty}, \right)$
-  where the family of equivalences are defined as follows
-  \[
-    a \nequal b \text{ iff }
-    \left\{
-      \begin{aligned}
-        a,b \in X \wedge a \nequal_X b \vee \\
-        a,b \in Y \wedge a \nequal_Y b
-      \end{aligned}
-    \right.
-  \]
-\end{lemma}
-\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
-%% Partial function from set
-\begin{lemma}[The c.o.f.e.\ of partial functions]
-  Given a set $S$ and a c.o.f.e $P$, the set of partial non-expansive functions $S \parfun P$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
-  \[
-    f \nequal[0] g \wedge f \nequal g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \nequal g(x)
-  \]
-\end{lemma}
-
 In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
-The future world relations is essentially a preorder structure that we would like to impose on our sets.
-We impose this structure on our sets by moving to \emph{preordered c.o.f.e.'s}.
+In order to have a future world relation that can be used on recursive instances of the world, we need to couple the future world relation to the sets we work with.
+The future world relations is a preorder structure, so we define a notion of \emph{preordered c.o.f.e.'s} which is simply a c.o.f.e.\ with a preorder that preserves the c.o.f.e.\ properties.
 % preordered cofe
@@
 \end{definition}
-The new structure we add to the set, the preorder, must be consistent with the existing structure which is why it must preserve limits.
+With the preordere c.o.f.e.\ definition, we can define the remainder constructs necessary to define the world we need.
+Any c.o.f.e.\ can trivially be lifted to a preordered c.o.f.e.\ where the ordering is the identity.
+
+The region invariants use monotone, non-expansive functions, so we need a c.o.f.e.\ to represent them.
+This c.o.f.e.\ is defined as follows.
+%% Monotone, non-expansive
+\begin{lemma}[The c.o.f.e. of monotone and non-expansive functions]
+  Given a c.o.f.e.
+$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future\right)$ where $\future$ is a preorder on $W$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ , define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ where
+  \[
+    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
+  \]
+\end{lemma}
+Next, we define the product preordered c.o.f.e.\
+%% Product
+\begin{lemma}[Product preordered c.o.f.e.]
+  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ define the product preordered c.o.f.e.\ as $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
+  where the family of equivalences are defined as follows
+  \[
+    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
+  \]
+  and the preorder $\future$ is defined as follows
+  \[
+    (x',y') \future (x,y) \text{ iff } x' \future x \wedge y' \future y
+  \]
+\end{lemma}
+The product c.o.f.e.\ is used to combine the sub-worlds into the complete world.
+The regions are also defined as a product, however, the preorder associated with the regions should correspond to the future region relation we sketched in \sectionname~\ref{subsubsec:ft-and-revocation}.
+Shared regions use the canonical ordering, but the canonical ordering does not suffice for spatial regions.
+Therefore, we define a preordered c.o.f.e.\ specifically for the construction of the spatial region tags.
+\begin{lemma}[Spatial region preordered c.o.f.e.]
+  Given a preordered c.o.f.e.\ $\left( H , \left( \nequal_H \right)_{n=0}^\infty , \future_H\right)$ the following is a preordered c.o.f.e.\ \[\left( \{\spatial,\spatial\} \times H + \{\pure\}, \nequal, \future \right)\]
+  where the family of equivalences $\left( \nequal \right)_{n=0}^\infty$ is the canonical family of equivalences and $\future$ is defined as the future region relation from \sectionname~\ref{subsubsec:ft-and-revocation}.
+\end{lemma}
+
+Similarly to the product preordered c.o.f.e. we have a sum preordered c.o.f.e.\ which is used to construct the heap sub-world.
+%% Sum
+\begin{lemma}[Sum preordered c.o.f.e.]
+  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ define the sum preordered c.o.f.e.\ as $\left(X \cup Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
+  where the family of equivalences are defined as follows
+  \[
+    a \nequal b \text{ iff }
+    \left\{
+      \begin{aligned}
+        a,b \in X \wedge a \nequal_X b \vee{} \\
+        a,b \in Y \wedge a \nequal_Y b
+      \end{aligned}
+    \right.
+  \]
+  and the preorder
+  \[
+    a \future b \text{ iff }
+    \left\{
+      \begin{aligned}
+        a,b \in X \wedge a \future_X b \vee{} \\
+        a,b \in Y \wedge a \future_Y b
+      \end{aligned}
+    \right.
+  \]
+\end{lemma}
+\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
+%% Partial function from set
+\begin{lemma}[The preordered c.o.f.e.\ of partial functions]
+  Given a set $S$ and a preordered c.o.f.e $\left( P, \left( \nequal_P \right)_{n=0}^{\infty}, \future_P \right)$, the preordered c.o.f.e.\ of partial non-expansive functions $\left(S \parfun P, \left( \nequal  \right)_{n=0}^\infty,\future\right)$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
+  \[
+     f \nequal g \text{ iff } 
+     \left\{
+       \arraycolsep=0pt \array{l}
+         n = 0 \vee{} \\
+         \dom(f) = \dom(g) \wedge \forall x \in \dom(f) \ldotp f(x) \nequal_P g(x)
+       \endarray
+     \right.
+  \]
+  and the canonical preorder
+  \[
+    f \future g \text{ iff } \dom(f) = \dom(g) \wedge \forall x \in \dom(f) \ldotp f(X) \future_P g(x)
+  \]
+\end{lemma}
 
@@
 % monotone functions
-\begin{lemma}
+\begin{definition}
   \label{def:monotone-preordered-cofe}
   Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty},\future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} ,\future_Y\right)$ be two preordered c.o.f.e.'s and $f$ a function from the set $X$ to the set $Y$.
-  We 
-  The function $f$ is
-  \begin{description}
-  \item[monotone] when for all $x, x' \in X$,
+  The function $f$ is \emph{monotone} when for all $x, x' \in X$,
 \[
@@
 \]
-  \end{description}
-\end{lemma}
+\end{definition}
+
 % TODO a preordered c.o.f.e. w/ our world?
@@
   \[
-    \blater\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right) = \left(X, \left(\nequal[n]'\right)_{n=0}^{\infty}, \future \right)
+    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
   \]
@@
   \[
-    x \nequal[0]' x' \wedge x \nequal[n]' x' \text{ iff } x \nequal[n+1] x'
+    x \nequal[0]_{\blater} x' \wedge x \nequal[n]_{\blater} x' \text{ iff } x \nequal[n+1]_X x'
   \]
@@
 The modality does not have to be placed directly on the self-reference.
-For instance, in the following it is placed on the world itself.
+For instance in the Theorem~\ref{thm:rec-dom-eq-sol}, it is placed on the world itself.
 The later placement does, however, affect what the logical relation definitions in Section~\ref{subsec:logical-relation} looks like, so in the end it is a matter of taste.
@@
   \begin{array}{l}
-    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
-    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
+    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg^2}) +\\
     \{\revoked\}
@@
 % TODO rewrite this introduction, add text between definitions (?), and reread this section.
-In order to express this, we will need to split the ownership of the world in disjoint parts.
-%% Disjoint union \oplus
-To this end, we introduce the operator $\oplus$ which is a disjoint union.
+The world serves multiple purposes as it is both a specification of memory contents as well as a specification of authority.
+This is best seen in the operators used to join worlds.
+First when we see the world as a memory specification, we have a pretty standard join $\uplus$ that simply requires that the two worlds do not use the same region names.
+\begin{definition}[World disjoint union $\uplus$]
+  Given worlds $W_1$, $W_2$, $W$
+  \[
+    W_1 \uplus W_2 = W
+    \text{ iff }
+    \begin{array}[t]{l}
+      \dom(\pwheap) = \dom(\pwheap[W_1]) \uplus \dom(\pwheap[W_2]) \tand \\
+      \dom(\pwfree) = \dom(\pwfree[W_1]) \uplus \dom(\pwfree[W_2]) \tand \\
+      \dom(\pwpriv) = \dom(\pwpriv[W_1]) \uplus \dom(\pwpriv[W_2]) \\
+    \end{array}
+  \]
+\end{definition}
+The $\uplus$ world join does not guarantee that the result is a sensible world with respect to authority.
+However, it does not guarantee that a join is a sensible memory specification.
+In \sectionname~\ref{subsubsec:mem-sat}, we define memory satisfaction which also acts as a well-formedness judgement.
+
+When we view the world as a specification of authority, then the world join need to respect the region ownership.
+That is, when we join the authority of two worlds, then the ownership of the two worlds should be completely separate.
+This is expressed by the $\oplus$ operator.
 \begin{definition}[$\oplus$, disjoint union of ownership]
@@
   \]
+  where $\oplus$ for regions is defined as
+\begin{align*}
+  (\pure,H,H_\sigma) \oplus (\pure,H,H_\sigma) =  & \; (\pure,H,H_\sigma) \\
+  (\spatial,H) \oplus (\spatial,H) =  & \; (\spatial,H) \\
+  \revoked \oplus \revoked = & \; \revoked \\
+  (\spatialo,H) \oplus (\spatial,H) = & \; (\spatial,H) \oplus (\spatialo,H)\\
+                                           =  & \; (\spatialo,H)
+\end{align*}
 \end{definition}
-
-\begin{definition}[World disjoint union $\uplus$]
-  Given worlds $W_1$, $W_2$, $W$
-  \[
-    W_1 \uplus W_2 = W
-    \text{ iff }
-    \begin{array}[t]{l}
-      \dom(\pwheap) = \dom(\pwheap[W_1]) \uplus \dom(\pwheap[W_2]) \tand \\
-      \dom(\pwfree) = \dom(\pwfree[W_1]) \uplus \dom(\pwfree[W_2]) \tand \\
-      \dom(\pwpriv) = \dom(\pwpriv[W_1]) \uplus \dom(\pwpriv[W_2]) \\
-    \end{array}
-  \]
-\end{definition}
-
-The two operators $\uplus$ and $\oplus$ are quite different.
-The difference is most clear in the treatment of pure regions: $\uplus$ allows both worlds to have the same pure region, while $\oplus$ forbids this.
-To understand this different treatment ($W_1 \uplus W_2$ and $W_1 \oplus W_2$), you should understand that the two are intended for different usages of worlds.
-The $W_1 \oplus W_2$ operator treats the worlds as specifications of authority: taking the disjoint union of worlds specifying non-exclusive ownership of a block of memory is allowed and produces a new world that also specifies non-exclusive ownership of world.
-The $W_1 \uplus W_2$  operator treats worlds as specifications of memory contents: taking the disjoint union of worlds specifying the presence of the same memory range is not allowed.
-The latter operator is used in the logical relation for components which specifies (among other things) that the world should specify the presence of the component's data memory.
+The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible, but when we use a world, we will always make sure that it is sensible by using memory satisfaction.
+
+The $\uplus$ operator is used in the logical relation for components (\sectionname~\ref{subsec:component-rel}) which specifies (among other things) that the world should specify the presence of the component's data memory.
 Linking two components then produces a new component with both components' data memory.
-The linked component is then valid in a world that has the combined memory presence specifications, not the combined authority.
-In other words, $\oplus$ specifies disjoint authority distribution, while $\uplus$ specifies disjoint memory allocation.
+The linked component is valid in a world that has the combined memory presence specifications, not the combined authority.
 
@@
 
-%% This both makes world a specification of authority as well as a specification of memory.
-
 \subsubsection{Memory satisfaction}
@@
 % Call stack
-The world also keeps track of the structure of the call stack, the allowed uses of designated seals, and linear capability authority, so these things also influence memory satisfaction.
+The world also keeps track of the structure of the call stack, the allowed uses of designated seals, and linear capability authority, so these things also influence the definition of memory satisfaction.
 % Well-formedness
-The memory satisfaction also acts as a well-formedness judgement.
-Our world definition allows regions to be overlapping in the sense that two regions may allow memory segments with the same addresses.
-We want the regions to be disjoint, so we easily can get the invariants of each memory segment.
+The world definition on its own allows the invariants imposed by regions to be overlapping.
+However, to be able to easily determine what invariants a memory segment must respect, we want the invariants of the regions to impose requirements on disjoint parts of the memory.
+The memory satisfaction relation makes sure that this is the case, so in a sense the memory satisfaction also acts as a well-formedness judgement for worlds.
 
 Memory satisfaction is split into four definitions.
-At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory, $\ms_S$, $\ms_\stk$, and $\stk$, from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
-The top memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
+At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory triple, $\ms_S$, $\ms_\stk$, and $\stk$, from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
+The top-level definition of memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
 It also splits the world in three to distribute the authority of the world.
-The three parts of $\ms_T$ are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrheap$, $\lrstk[]$, and $\lrfree[]$, respectively. 
+The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrheap$, $\lrstk[]$, and $\lrfree[]$, respectively. 
 The $\lrheap$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
-The $\lrheap$ relation also ensures that the seal invariants associated with each regions do not put invariants on the same seals.
+The $\lrheap$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
 The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
-The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the stack frames have the correct order relative to each other as well as the base address of the stack.
+The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the placement in memory of the stack frames have the correct order relative to each other and the base address of the stack.
 Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
-$\lrfree[]$ also makes sure that that base address of the stack is actually part of the stack.
+$\lrfree[]$ also makes sure that the base address of the stack is actually part of the free stack.
 \begin{definition}[Heap relation]
@@
       \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
-      \quad \forall r \in \dom(\activeReg{\pwheap})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
+      \quad \forall r \in \dom(\erase{\pwheap}{\pure})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
     \end{array}
@@
 
-To a large extend, the way $\lrheap$ is defined is pretty standard.
-$\lrheap$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into pairs each satisfying the invariant of a region.
-The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction also partition the authority of the world.
+To a large extend, the definition of $\lrheap$ is pretty standard.
+$\lrheap$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
+The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction partitions the authority of the world using $\oplus$.
 Each of the memory segment pairs must be in the region invariant with respect to a specific world partition which makes sure that uniqueness of linearity of capabilities is respected.
@@
 The heap sub-world contains all seal invariants.
-Similar to memory segments, only one seal invariant should impose restrictions on a seal which $\lrheap$ makes sure is the case.
+Similar to memory segments, only one seal invariant should impose restrictions on a seal, which $\lrheap$ makes sure is the case.
 \begin{definition}[Free stack relation]
@@
 \end{definition}
-The free stack relation $\lrfree[]$ is in most regards like the heap relation.
-The free stack relation, partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under the
+The free stack relation $\lrfree[]$ is in most regards like the heap relation, $\lrheap{}$.
+The free stack relation, partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under part of the world.
+For \stktokens{} to work, it should always work on the same stack.
+As discussed in \sectionname~\ref{sec:stktokens-explained}, we make sure that it is always the same stack by requiring the address $\stkb$ to be the ''top'' address of the free stack address space.
+As the free stack relation relates the stack of \srccm{} with the memory that represents the stack on \trgcm{}, it makes sure that $\stkb$ is the top address of the free stack address space.
 \begin{definition}[Stack relation]
@@
 The stack relation $\lrstk[]$ is similar to the heap relation in some ways.
-$\lrstk[]$ also partitions the memory but only the \trgcm{} as the \srccm{} partitions are given as the stack frames, and it also partitions the authority of the world, so it can relate the stack frames such that they respect linearity.
-The stack on \srccm{} represents the call stack which means that each stack frame represents a call and its local data.
-The operational semantics of \trgcm{} does not enforce the same stack structure, so we make sure that a \srccm{} stack is only related to \trgcm{} stack that has the same structure.
-That is for a memory segment that represents a stack frame, all the addresses of memory frames lower in the stack should have strictly smaller memory addresses .
-Further, the stack frames should be in the part of the memory we agree to be the stack which means that the addresses should be less than $\stkb$.
+The $\lrstk[]$ relation also partitions the \trgcm{} memory but only the \trgcm{} as the \srccm{} partitions are given as the stack frames.
+The stack relation also partitions the authority of the world, so it can relate the stack frames in a way that respect linearity.
+The stack on \srccm{} represents the call stack which means that each stack frame corresponds to a call and its local data.
+The operational semantics of \trgcm{} does not have a built-in stack, so we emulate it by requiring that a stack like data structure resides in \trgcm{} memory.
+That is for a memory segment that represents a stack frame, all the addresses of memory frames lower in the stack should have strictly smaller memory addresses.
+Further, the stack frames should be in the part of the memory we agree to be the stack which means that the addresses should be smaller than $\stkb$.
 Informally, this just means that the stack should be laid out in memory as a downwards growing stack with no addresses below $\stkb$.
+
 % TODO add figure that illustrates this.
@@
 For this reason, $\lrstk[]$ requires every stack frame to be non-empty.
-Each stack frame in the \srccm{} stack contains an old program pointer.
-Each of the old program counters on the stack should correspond to the old program counter recorded in the region of the world associated with the stack frame.
-To achieve this, the partitioning function $R_\ms$ also records an $\opc$ for each region, and this $\opc$ should establish the link the region and the stack frame.
-
-In order to tie Definition~\ref{def:heap-rel},~\ref{def:free-stack-rel},~and~\ref{def:stack-rel}, we define memory satisfaction.
-Memory satisfaction defines when a \srccm{} memory , consisting of a heap, a stack, and a free stack, relates to a \trgcm{} memory under a world.
+Note that each stack frame corresponds to a trusted call.
+Untrusted calls are not protected which means that untrusted stack frames reside in the free stack memory (this does not mean that the untrusted stack frames are necessarily unprotected; it only means that they are not natively protected by \stktokens{}).
+This means that the protected stack frames are not necessarily packed tightly in memory, and the memory in between is part of the free stack.
+% TODO illustrate the protected stack ?
+Each stack frame in the \srccm{} stack contains an old program pointer which corresponds to the old program counter recorded in the region of the world associated with the stack frame.
+To achieve this, the partition function $R_\ms$ also records an $\opc$ for each region, and this $\opc$ should establish the link between the region and the stack frame.
+
+In order to tie Definitions~\ref{def:heap-rel},~\ref{def:free-stack-rel},~and~\ref{def:stack-rel} together, we define memory satisfaction.
+Memory satisfaction defines when a \srccm{} memory, consisting of a heap, a stack, and a free stack, relates to a \trgcm{} memory under a world.
 \begin{definition}[Memory satisfaction]
@@
 Our configurations describe the complete machine state, but we may only be interested in the invariants on part of it.
-The frame allows us to basically ignore the part of the memory that won't affect the computation.
+The frame allows us to ignore the part of the memory that won't affect the computation.
 Just like the previous memory relations, the world is split in three to make sure that linearity is respected.
@@
 \stktokens{} requires the stack to not be adjacent to heap or code memory.
-This is enforced in the memory satisfaction by requiring that the addresses next to the memory are in the frame.
+This is enforced in the memory satisfaction by requiring that the addresses adjacent to the memory are in the frame.
 
@@
   \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
-  \npair{(\reg_S,\reg_T)}  &  \lrrtrusted(W) & register files & \\
   \npair{\Phi_S,\Phi_T}  &  \lro & execution configurations & \\
@@
 \end{tabular}\\
-These relations are defined using a set of mutually recursive equations, with cyclicity resolved through another use of step-indexing.
 In Section~\ref{subsubsec:mem-sat}, we already defined memory satisfaction, the relation for memories.
 In the following, we define each of the remaining relations and give some intuition about the definitions.
-The logical relation we define ends up as a recursive definition.
-The circularity is resolved by a step-index in the definitions, but the circularity also poses a chicken and egg kinda situation with respect to the order in which the definitions of the relations should be presented.
+The logical relation we define ends up as a cyclic definition.
+The circularity is resolved by another use of step-indexing in the definitions, but the circularity also poses a chicken and egg kinda situation with respect to the order in which the definitions of the relations should be presented.
 There is no canonical way of presenting the logical relation as we are bound to make forward references.
-For this reason, we suggest making a cursorily first read through followed by a more thorough read.
+For this reason, we suggest making a cursorily first read through to get an overview followed by a more thorough read.
+% TODO something about relation naming?
 
 \subsubsection{Observation relation}
-% TODO write section
-\[
-  \lro[\preceq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+The observation relation defines what machine configurations have related and permissible observable effects.
+Generally speaking, an observation relation captures the property we want to prove.
+% TODO double check "co-termination" (equitermination?)
+Ultimately, we want to prove a full-abstraction theorem which is defined in terms of contextual equivalence for components that in turn is defined as co-termination in any context.
+This means that the observation relation should capture co-termination.
+
+So far, we have talked about the logical relation as though we define one.
+However, we actually define two logical approximations that only differ in the observation relation.
+We define a \srccm{} configuration to logically approximate a \trgcm{} configuration when the halting termination of the \srccm{} configuration implies the halting termination of the \trgcm{} configuration.
+This also means that \srccm{} configurations that terminates by failing termination are related to any \trgcm{} configuration.
+Intuitively, this is because the $\failed$ configuration signals that there was an attempt to break the guarantees of the capability machine.
+For instance, a piece of code could have attempted to read from a part of memory it does not have access to, or a callee could have attempted to return out of order.
+In both cases, we haven't defined a way to recover from such attempts to break the guarantees, so we are content with failure.
+\begin{multline*}
+  \lro[\preceq,(\ta,\stkb,\_,\_)] \defeq \\\left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
@@
 \right\}
-\]
-\[
-  \lro[\succeq,(\ta,\stkb,\_,\_)] = \left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
+\end{multline*}
+The step-indexing plays a role here because we are only interested in \srccm{} configurations that terminate in $n$ or fewer steps.
+However, if the \srccm{} configuration terminates successfully in $n$ steps, then the \trgcm{} configuration should just terminate in any number of step (possibly more than $n$ steps).
+For the most part, it would make sense to require the \trgcm{} configuration to terminate in the same amount of steps as the \srccm{} configuration as they run in lockstep for most of the computation.
+However, when it comes to calls and returns, the two configurations stop running in lockstep.
+The \srccm{} configuration handles calls and returns in one step whereas \trgcm{} configurations need to execute each instruction of the call preparation as well as the return code.
+
+We define a \trgcm{} configuration to approximate a \srccm{} configuration in a dual way to the above.
+\begin{multline*}
+  \lro[\succeq,(\ta,\stkb,\_,\_)] \defeq \\\left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
     \begin{array}{l}
@@
 \right\}
-\]
+\end{multline*}
+The remainder of our logical relation will be the same for both $\preceq$ and $\succeq$, so we will write $\square$ instead of the approximation.
 
@@
 \citet{skorstengaard_reasoning_2017} defines a logical relation that can be seen as a notion of capability safety.
-When they define their value relation, they define based on the question ``What is the most we can allow an adversary to do with this word without breaking memory guarantees?''
+When they define their value relation, they define based on the question ``What is the most an adversary can be allowed to do with this word without breaking memory invariants?''
 This allows them to use the logical relation to reason about arbitrary (untrusted) programs.
@@
 In our setting, a trusted program is a well-formed, reasonable program that follows the \stktokens{} calling convention, and an untrusted program is an arbitrary well-formed program.
-In order for \stktokens{} to work, it needs access to return seals, but we cannot allow untrusted programs access to the return seals.
-A value relation based on what an adversary could do should permit return seals, so such a relation cannot be used to reason about trusted programs.
+In order for a trusted program to use \stktokens{}, it needs access to return seals, but we cannot allow untrusted programs access to the return seals.
+A value relation based on what it is safe for an adversary to have should prohibit return seals, so such a relation cannot be used to reason about trusted programs.
 For this reason, we define two value relations a trusted $\lrvtrusted$ and an untrusted $\lrv$.
@@
 
-In this section, we will from time to time refer to safety of a capability or a word.
+From time to time in this section, we will refer to safety of a capability or a word.
 In some sense, our logical relation actually ends up as the definition of safety, so when we refer to a capability as \emph{safe} it is in an informal sense where it means that the capability cannot be used break memory invariants.
@@
 Note that the return pointers of \srccm{} are not related to anything as it is never safe for any program, trusted or not, to have them.
-The \srccm{} return pointers should only occur under a seal, and they should only be used to jump to in which case the \srccm{} semantics transforms them to the capabilities they correspond to.
+The \srccm{} return pointers should only occur under a return seal, and they should only be used in a jump in which case the \srccm{} semantics transforms them to the capabilities they correspond to.
 
@@
 
-We define a $\pure$, $\spatial$, and $\spatialo$ standard. They all have the same invariant which is defined as follows:
+We define a $\pure$, $\spatial$, and $\spatialo$ standard region. They all have the same invariant which is defined as follows:
 \[
@@
       \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
-      \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\ % TODO there should be a later on the right side of the equality?
+% There should be a later on the right side of the equality? Lau: To myself, on second thought we leave that implicit. This should be fine.
+      \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\ 
       \quad \forall \aaddr \in A \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
@@
 \]
-Note that the standard $\pure$ region puts no requirements on seals.
+Note that the standard $\pure$ region has an empty seal invariant and thus puts no requirements on seals.
 
 Sometimes we need to know that the contents of a memory segment stays the same.
-For instance, the contents of encapsulated stack frames don't change which we need to be able to rely on.
+For instance, the contents of encapsulated stack frames do not change which we need to be able to rely on.
 To express this, we define a \emph{static region}.
@@
 The region also requires the static memory to contain words from the untrusted value relation.
-Which just means that the stack shouldn't be used for return seals, closure seals, and code pointers for trusted code.
+This means that the stack should not be used to store return seals, closure seals, and code pointers for trusted code.
 With the memory invariant, we define the static region as follows:
@@
 \end{multline*}
-The code region is a lot more restrictive that the standard region.
+The code region is a lot more restrictive than the standard region.
 It basically only allows one memory segment, namely $\code$ padded with zeroes to make sure that two capabilities cannot be spliced to cause unintended control-flow.
@@
 This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
-This is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then it is trusted.
-On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then it is untrusted.
+That is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then the component is trusted.
+On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then the component is untrusted.
 In either case, the words should be in the value relation with respect to the $\purePart{}$ of the world which means that the code memory cannot contain linear capabilities.
-% TODO promote $\purePart{}$ from appendix to paper - used a few times.
 
 \stktokens{} rely on proper seal usage to guarantee well-bracketed control-flow and local state encapsulation.
-This means that components must use return and closure seals for their intended purpose.
+This means that components must use return and closure seals for their intended purpose for \stktokens{} to work.
 The code region has a seal invariant $H^\mathrm{code,\square}_\sigma$ to guarantee that the return and closure seals of the region are used correctly.
@@
 The return seals $\sigrets$ in a code region should only be used to seal return pointers.
-That is on \srccm{} they should only be used to seal $\retptrc{}$ and $\retptrd{}$.
+That is on \srccm{}, the return seals should only be used to seal $\retptrc{}$ and $\retptrd{}$.
 \begin{figure}
@@
 % Code return pointer
-If we allowed any $\retptrc$ to be sealed, then we could not be sure that it comes from a call even though it should only be possible to get a return pointer from a call.
+If we allowed any $\retptrc$ to be sealed, then we could not be sure that the $\retptrc$ came from a call even though it should only be possible to get a return pointer from a call.
 For this reason, we require that the \srccm{} return pointer actually points to the first address after a call.
-For a \trgcm{} capability related to a \srccm{} code return pointer, we do not require that it points to the first address after the call because it needs to execute the return instructions first.
+For a \trgcm{} capability related to a \srccm{} code return pointer, we require it to point to the first address of the return code, not the first address after the call, as the return instructions must be executed.
 
@@
 That is, there should be a static region with the contents of the stack frame.
-The fact that it is static means that the contents will remain the same.
+The fact that it is static signifies that the contents will remain the same.
 The region that governs the stack frame must come from the private world which means that it is paired with a return address.
-We require that the return address actually corresponds to a return address for a call in the code.
+The return address should correspond to an actual return address of a call in $\code$.
 
@@
 For untrusted components (components with their code address space disjoint from the trusted address space), we allow everything in the untrusted value relation to be sealed.
-Intuitively, untrusted components are assumed to have access to words from the untrusted value relation, but we cannot know how the words are used, so we need to assume that they may be sealed.
+Intuitively, untrusted components are assumed to have access to words from the untrusted value relation, and we cannot know how the words are used, so we need to assume that an untrusted component may seal untrusted words.
 Trusted components only use closure seals for sealed capability pairs that represent actual closures.
@@
 For this reason, the seal invariant allows executable capabilities from the trusted value relation to be sealed with a closure seal.
-When it comes to the data capability of a closure, we just require that it comes from the untrusted value relation because the trusted value relation contains nothing that makes sense to seal as the data capability (we return to the specific contents of the two value relations in a later paragraph).
+
+When it comes to the data capability of a closure, we just require that it comes from the untrusted value relation because the trusted value relation contains nothing that makes sense to seal as the data capability
+% TODO be more precise, what later paragraph means.
+(we return to the specific contents of the two value relations in a later paragraph).
 
@@
 The code region is $\pure$ because it needs to contain a seal invatiant and because we assume that code pointers are non-linear.
+
+% TODO in the above, add something about the untrusted case where the seals should nut be in the "global" seals because the global seals are the one that we globally agree to be trusted.
 
@@
 All the permission based condition we define project the regions that the capability may address from the world.
-The following $\addressable$ function takes care of the projection:
+The addressable $\addressable{}$ function takes care of the projection:
 \[
@@
 % each of the regions upper bounded by std region
-The main purpose of $\readCond{}$ is to make sure that only safe words can be read from the memory the read capability governs.
-This is done by putting an upper bound on what requirements a region can impose on the memory segments governed by the capability.
+The main purpose of $\readCond{}$ is to make sure that only safe words can be read from the memory governed by a read capability..
+This is done by putting an upper bound on what requirements an invariant can impose on the memory segments governed by the capability.
 In particular a region that governs the memory a read capability has access to can at most allow safe values to be read.
-Without this requirement, a read capability could potentially be used to break memory invariants if were used to read capabilities that has the authority to break memory invariants.
-The condition is defined as follows
+Without this requirement, a read capability could potentially be used to break memory invariants if it were used to read capabilities that has the authority to break memory invariants.
+The read condition is defined as follows
 \[
@@
 The $\readCond{}$ is compatible with all the operations that can be performed on capabilities.
-This means that if two capabilities, for which $\readCond{}$ holds, are spliced together, we should be able to establish that $\readCond{}$ holds for the resulting capability.
+This means that if two capabilities, for which $\readCond{}$ holds, are spliced together, we can establish that $\readCond{}$ holds for the resulting capability.
 To support this, we require the presence of a set of regions $S$ that governs the addresses the capability has authority over rather than just a single region.
-If we need to establish the $\readCond{}$ after a splice, we can simply use union of the regions that witnessed the $\readCond{}$ of the two individual capabilities.
+If we need to establish the $\readCond{}$ after a splice, we can simply use the union of the regions that witnessed the $\readCond{}$ of the two individual capabilities.
 We also need to support splitting which is no problem for non-linear capabilities as the same $\pure$ region can be used to establish the $\readCond{}$ for multiple non-linear capabilities.
@@
 This is why, $\readCond{}$ requires all regions to only govern one address when the capability is linear.
-This means that after a split, the authority of the regions for the bottom half of the split result can go to one capability and the rest can go to the top half.
-
+This means that after a split, the authority of the regions for the bottom half of the split can go to one capability and the remaining regions can go to the top half.
+
+% TODO the following paragraph does not read well. Rewrite.
 The upper bound on requirements that can be imposed on the memory the read capability has access to is specified on all of the regions that govern the memory.
@@
 The question is, what can we safely allow to be written to memory without any memory invariants being broken.
-It may come as a surprise that the answer to this is anything - even words that are unsafe.
-The intuition behind this is that if you managed to write something that can break memory invariants, it would not be possible to read it back again as you need read permission for that, and if the capability had read permission, then $\readCond{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants, after all if one was obtained, memory invariants could be broken.
+The answer to this is anything - even words that are unsafe.
+Say, you manage to write something that can break memory invariants, then it would not be possible to read it back again as write permission, generally speaking, does not entail read permission.
+Iff the capability had read permission, then $\readCond{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants.
+After all, if such a capability was obtained, memory invariants could be broken.
   However, the $\writeCond{}$ tries to capture the essence of safety and in principle it is safe to write an unsafe capability that cannot be read back.}.
+It should always be possible to write safe values, so we impose this as a lower bound.
 We do, however, impose a bound on what should be permitted to be written and that is a lower bound because it should at least be allowed to write anything that is safe.
@@
 Before we get to the definition of $\writeCond{}$, we need to introduce one auxiliary definition.
-When a write capability makes a change to memory, the change must be permitted by the regions that govern the memory it has authority over for it to be safe.
-Given a write capability it should be possible to transform the memory from one allowed by the region to another also allowed by the region.
+When a write capability makes a change to memory, the change must be permitted by the regions that govern the memory it has authority over for it to not break the invariant and thus be safe.
+Given a write capability, it should be possible to transform the memory from one allowed by the region to another also allowed by the region.
 However, only one memory address can be updated at a time, so in order to make such a transformation one has to update one address at a time.
@@
 Note that we do not define $\execCond{}$ and $\xReadCond{}$ for the stack because the stack is not executable.
+
 \paragraph{The untrusted value relation}
 \label{par:untrusted-val-rel}
-% TODO write section
+The untrusted value relation $\lrv$ relates all the words that untrusted components can safely posses.
+That is words that cannot be used to break any memory invariants.
+The relation is displayed in \figurename~\ref{fig:untrusted-val-rel}.
+
+The untrusted value relation has five cases: data, capabilities, stack pointers, sealed capabilities, seal sets, and stack pointers.
+In the following, we will give some intuition about why it is safe to give these words to untrusted code as well motivate the conditions they are safe under.
+
+The first case is data.
+It is always safe to posses data as data is just integers that grant no authority.
+Further, unlike capabilities, it is always possible to construct a new integer with the move instruction.
+
+The next case is capabilities that do not have a special representation on \srccm{} unlike stack pointers and return pointers.
+For two capabilities to be related, they should be syntactically equal. That is, they should have same range of authority, linearity and so one.
+Generally speaking, untrusted components should not have direct access to a trusted components code.
+For this reason, we require that capabilities must have a range of authority outside the trusted address space if they are to be related.
+The safety of a capability also depends on the world and whether the capability can be used to break the memory invariants of the world.
+For instance, if a capability has read-permission, then it should only by able to read safe values with it.
+This condition and conditions for the other permissions are captured by the permission based conditions, so we use them to express the necessary conditions.
+That is, if a capability has read permission, then $\readCond{}$ must be satisfied, if it has writer permission, then the $\writeCond{}$ must be satisfied, and if it has execute permission, then $\execCond{}$ and $\xReadCond{}$ must be satisfied.
+If the capability has execute permission, then it must also be a non-linear capability.
+Finally, the capability cannot have read/write/execute permission because that would break the write-XOR-execute assumption.
+
+Stack pointers on \srccm{} are represented with the special token $\stkptr{\perm,\baddr,\eaddr,\aaddr}$.
+The corresponding capability on \trgcm{} is a linear capability with the same permission and addresses.
+We assume that the stack is non-executable, so the permission for a stack pointer cannot have execute permission.
+Similarly to the normal capabilities, we use the permission based conditions for the stack to ensure that the stack capability is safe to use.
+
+% Encapsulation
+A sealed capability encapsulates the authority of the underlying capability, and the authority is only released when the sealed capability is used in an $\txjmp{}{}$.
+The $\xjmp{}{}$ takes a pair of sealed capabilities, so the authority of a sealed capability depends on what other sealed capabilities it might be used with.
+As discussed, closure seals and return seals must be used in specific ways, so we have defined a seal invariant that dictates what may be sealed with a given seal.
+In order for a pair of \srccm{} and \trgcm{} seals to be in the unsafe value relation, they must be sealed with the same seal $\sigma$, and they must be related by the seal invariant $\sigma$.
+% Sealed capability only useful with pairing. Sealed caps must be permitted.
+Further, if we take any sealed capabilities related under $\seal$ in the seal invariant, then they should produce a safe result.
+We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrexj$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
+% Linear and non-linear case. Affects the future world we consider.
+It should be safe to use the sealed capability both now and later.
+This means that it should be safe in the world $W$, but also in future worlds $W' \future W$.
+However, this only goes for non-linear words.
+For linear capabilities, it should only be in future worlds of the pure part of $W$.
+% TODO write why we have the two cases. (it allows us to have the "other" capability by the one we consider - and it feels like the only reason)
+
+For sets of seals to be related related in the untrusted relation they must be syntactically equal.
+Further, the seals in the set should be disjoint from the return seals and trusted closure seals ($\gsigrets$ and $\gsigcloss$ because trusted code relies on having the sole access to them.
+We do not know what an adversary may seal or what seal they may use, so, for every seal in the seal set, we require the seal invariant to be essentially equal to the untrusted value relation.
+
+% TODO add approx square to relation, add gc
+\begin{figure}
+  \centering
   \[
@@
 %
-    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
-    \begin{array}{l}
-      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
-      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
-      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
-
-      % This case was never used (?) so it is not necessary.
-      % \wedge \\
-      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
-    \end{array}
-    \right\}\cup\\
-    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
-    \; \middle| \;
-    \begin{array}{l}
-      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
-      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
-    \end{array}
-    \right\} \cup \\
-    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
-    \begin{array}{l}
-      \begin{array}{r l l}
-        \perm \not\in \{\rx,\rwx\} \wedge\\
-        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
-        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
-        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
-      \end{array}
-    \end{array}
-    \right\} \cup \\
-%
     \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
     \begin{array}{l}
-      [b,e] \mathrel{\#} \ta \tand\\
+      [b,e] \mathrel{\#} \ta \wedge\\
       \begin{array}{r l l }
@@
     \end{array}
+    \right\}\cup \\
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
+    \begin{array}{l}
+      \begin{array}{r l l}
+        \perm \not\in \{\rx,\rwx\} \wedge\\
+        % \perm = \noperm & \Rightarrow & \npair{(\linear,\baddr,\eaddr)} \in \lrp(W) \wedge \\
+        \perm \in \readAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackReadCond{W} \wedge \\
+        \perm \in \writeAllowed{} &\Rightarrow& \npair{[\baddr,\eaddr]} \in \stackWriteCond{W}
+      \end{array}
+    \end{array}
+    \right\} \cup \\
+%
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
+    \begin{array}{l}
+      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
+      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
+      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
+      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
+      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
+      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
+
+      % This case was never used (?) so it is not necessary.
+      % \wedge \\
+      % \quad \npair[n']{\src{\vsc_S'},\src{\vsc_S},\vsc_T',\vsc_T} \in \lrexj(W' \oplus W_o)
+    \end{array}
+    \right\}\cup\\
+    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma_\baddr,\sigma_\eaddr,\sigma}},\\ \seal{\sigma_\baddr,\sigma_\eaddr,\sigma} \endarray \right)} 
+    \; \middle| \;
+    \begin{array}{l}
+      [\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss) \tand\\
+      \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
+      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+    \end{array}
     \right\} 
@@
 \]
+\caption{The untrusted value relation relates all the words on \srccm{} to all the words on \trgcm{} that are safe for non-trusted components to posses.}
+\label{fig:untrusted-val-rel}
+\end{figure}
 
@@
 % TODO write section
+The trusted value relation $\lrvtrusted$ relates everything safe for a trusted component to have without breaking memory invariants.
+For the most part, we allow them to contain the same words as the untrusted components, but we also need to allow them to have seal sets with trusted closure seals and return seals which we cannot allow untrusted components to have.
+Further, we need to allow trusted components to have capabilities for the trusted code which, again, is something that we cannot allow untrusted components to have.
+The untrusted value relation is defined in Figure~\ref{fig:trusted-val-rel}.
+
+% Trusted seals
+The trusted closure seals and return seals serve a specific purpose, namely they must be used for return pointers and closures.
+To make sure this is the case, there must be a code region in the world that governs the code.
+The code region contains a seal invariant that makes sure that the seals are only used for their intended purpose.
+This is why the trusted value relation only relates seal sets of trusted closure seals and return seals when the world contains an appropriate code region.
+
+% Code pointers for trusted code
+Two capabilities are related as trusted code pointers if they are non-linear, has a permission derivable from read-execute, and has a range of authority within the trusted address space, $\ta$.
+Further, we need to know that they are actually capability which is why we require the $\xReadCond{}$ to be satisfied.
+This makes sure that the region that governs the memory the capability points to is a code region.
+Note that even though the capability has read permission, we do not require the read condition to hold.
+The code memory contains trusted closure seals and return seals that we cannot let untrusted code have, so the read condition would not hold.
+It is, however, fine to let trusted code have access to such seals because we expect it to treat the seals reasonably.
+\begin{figure}
+  \centering
   \[
@@
 
+\caption{The trusted value relation $\lrvtrusted$ relates all the words that are safe for trusted components to contain.
+A trusted component may contain untrusted words (\figurename~\ref{fig:untrusted-val-rel}), return seals and trusted closure seals, and code pointers for trusted code.
+}
+\label{fig:trusted-val-rel}
+\end{figure}
 \subsubsection{Register file relation}
-In Figure~\ref{fig:obs-rel}, we show sketches of the register file relation $\lrr$, the relation between $\tjmp{}$ targets $\lre(W)$ and the observation relation $\lro$.
-The trusted/untrusted register file relation simply requires that all registers except $\pcreg$ are in the corresponding value relation (in a compatible world partition).
-Two execution configurations are in the observation relation $\lro$ if one terminates whenever the other does\footnote{The actual definition in the \cite{technical_report} is complicated a bit by step-indexing and the fact that we actually use two separate observation relations for left- and right-approximation.}.
-The $\lre(W)$ relation then includes any two words which can be plugged into related register files and memories (in compatible worlds), to obtain execution configurations in the observation relation.
-
-\[
-  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
+The register file relation relates \srccm{} register files to \trgcm{} register files.
+Intuitively, two register files are related when they only contain safe words, i.e.\ words from the value relation.
+This raises the question ``which value relation''?
+We only use the register file relation to relate register files for components we do not trust, so the answer is the untrusted value relation.
+The definition of the value relation is straightforward.
+It distributes the authority of the world among the registers and require each of the registers to contain a safe word with respect to the authority it is given.
+The register file never takes into account the $\pcreg$ and it can leave out further registers.
+We use this to not relate register content that we are going to overwrite with new values.
+We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$.
+That is, if we do not need to exclude extra registers, then we simply omit that argument.
+The register file relation is defined in \figurename~\ref{fig:reg-file-relation}.
+
+\begin{figure}
+  \centering
+  \[
+  \lrr(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
     \begin{array}{l}
@@
       \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
-      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
+      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
     \end{array}
@@
 \]
-We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
-
-\begin{lemma}[Untrusted is trusted]
-  \label{lem:untrusted-supset-trust}
-  \begin{itemize}
-  \item $\lrvtrusted(W) \supseteq \lrv(W)$
-  \item $\lrrtrusted(W) \supseteq \lrr(W)$
-  \end{itemize}
-\end{lemma}
+\caption{The register file relation relates register files. Two register files are related when their content is related.}
+\label{fig:reg-file-rel}
+\end{figure}
 
@@
 % TODO write section
-% Two expression relations: one for sealed code-data pairs being jumped to and one for capabilities being jumped to in the regular way.
-% The argument for having one relation relate pairs of pairs of capabilities and the other relate pairs of capabilities is that that is how xjump and regular jumps work: xjump takes pairs while regular jumps take single capabilities.
+The expression relation $\lre$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e. construct configurations in the observation relation.
+The $\lre$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control-flow control-flow during execution when a $\tjmp{}$ instruction is executed.
+In the setting of \srccm{} and \trgcm{}, we can also use sealed capabilities to change the control flow by using the $\txjmp{}{}$ instruction.
+The $\txjmp{}{}$ updates the $\pcreg$ register and the $\rdata$ register, however, the $\lre$ relation only updates the $\pc$-register, so we cannot use $\lre$ to reason about sealed capabilities.
+Instead, we define the relation $\lrexj$ which relates two pairs of capabilities when they are safe to use with the $\txjmp{}{}$ instruction.
+
+In this case, related executions means that the observable effect of the executions should be permissible.
+The permissible observations are defined by the observation relation, so we define the expression relation in terms of the observation relation.
+However, the observation relation relates configurations, not capabilities.
+We lift the capabilities to configurations simply by plugging the two capabilities into the $\pcreg$-register of two configurations.
+We cannot pick arbitrary configurations because an arbitrary configuration may contain words that can be used to break memory invariants and create observable effects that we do not permit.
+Instead, we need to pick configurations made out of related components, i.e.\ related register files and related memories that respect linearity.
+The type of execution captured by $\lre$ corresponds to a normal jump.
+When a $\tjmp r$ instruction is interpreted, the $\pcreg$-register is replaced with the contents of register $r$, i.e. the current configuration is plugged with a new pc.
+The $\lre$ relation is defined in \figurename~\ref{fig:expr-rels}
+
+The $\lrexj$ relation looks very much like the $\lre$ relation.
+It takes related memories and register files (ignoring the $\rdata$ register) and combines them into two configurations.
+Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lrexj$ relation.
+The $\lrexj$ relation is defined in \figurename~\ref{fig:expr-rels}.
+\begin{figure}
+  \centering
   \begin{align*}
-  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+% TODO add approx square
+  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
+    \begin{array}{l}
+      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
+      \quad \forall W_\lrrs , W_\lrm \ldotp \\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \wedge\\
+      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
+      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
+      \qquad\Phi_T = (\ms_T,\reg_T)\\
+      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
+      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
+      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+    \end{array}
+    \right\}
+\\  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
     \begin{array}{l}
@@
     \end{array}
-    \right\}\\
-  \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
-    \begin{array}{l}
-      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
-      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
-      \qquad\Phi_T = (\ms_T,\reg_T)\\
-      \qquad\Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}}\\
-      \qquad W \oplus W_\lrrs \oplus W_\lrm\\
-      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
-    \end{array}
     \right\}
 \end{align*}
-\end{jversion}
+
+\caption{The expression relation relates capabilities capabilities that can safely be used for execution.
+The xjmp expression relation can be used to relate capabilities that are safe as sealed capabilties.}
+\label{fig:expr-rels}
+\end{figure}
+
 \subsection{Fundamental Theorem}
-% TODO rewrite section
 % Include the actual FTLR
@@
 The name indicates that it is an instance of a general pattern in logical relations proofs, but is otherwise unimportant.
-\begin{lemma}[FTLR (roughly)]
+\begin{theorem}[FTLR]
   \label{thm:ftlr}
-  If $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and $\pwheap(r) = \codereg{\sigrets,\sigcloss,\mscode}$, and either $[\baddr,\eaddr] \subseteq \ta$ and $\mscode$ behaves reasonably (see Section~\ref{sec:well-form-reas}) or
-$[\baddr,\eaddr] \mathrel{\#} \ta$,
+  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
+  If
+  \begin{itemize}
+  \item $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
+  \end{itemize}
+  and one of the following sets of requirements holds:
+  \begin{itemize}
+  \item $[\baddr,\eaddr] \subseteq \ta$ and
+ $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps (see Section~\ref{sec:well-form-reas}).
+  \item $[\baddr,\eaddr] \mathrel{\#} \ta$
+  \end{itemize}
   then
@@
   \]
-\end{lemma}
-
+\end{theorem}
 Roughly speaking, this lemma says that under certain conditions, executing any executable capability under \srccm{} and \trgcm{} semantics will produce the same observable behavior.
@@
 \subsection{Related Components}
-% TODO write section
+\label{subsec:component-rel}
+In order to show full abstraction (Theorem~\ref{thm:full-abstraction}), we need not only to relate the words on \srccm{} with words on \trgcm{} we also need to relate \srccm{} components with \trgcm{} components.
+Specifically, we say that two components are related when they are syntactically equal after all a \srccm{} component is in some sense the same as a \trgcm{} as we only see the difference during execution when a call happens and when we lift a component to a configuration which means that we need to introduce a stack pointer.
+However, we cannot take arbitrary components as they could potentially break memory invariants.
+For related components, we require that if the imports are satisfied with related words, then the resulting memory should be safe.
+Further, related components must have safe exports.
+This roughly describes when a base components are related.
+A component with a main is related when the base component is related and the mains capabilities are in the exports.
+
+\begin{definition}[Component relation]
+  \label{def:component-rel}
 \begin{align*}
@@
 \end{align*}
+\end{definition}
 % TODO include central lemmas?
@@
 \subsection{Related Execution Configuration}
-% TODO write section
-\[
-  \lrec(W) = \left\{ \npair{\left(
-        \arraycolsep=0pt\array{l}
-        (\ms_S,\reg_S,\stk,\ms_\stk),\\
-        (\ms_T,\reg_T)\endarray\right)} \middle|
+The full abstraction theorem (Theorem~\ref{thm:full-abstraction}) is stated in terms of contextual equivalence.
+Contextual equivalence (Definition~\ref{def:contextual-equivalence}) plugs two components with a context and requires equitermination of the resulting executable configurations.
+This means that we need to lift relatedness one step further, namely to the level of execution configurations.
+To this end, we define $\lrec$.
+\begin{definition}[Related execution configuration]
+\label{def:exec-conf-rel}
+  \[
+  \lrec(W) = \left\{
+    \begin{multlined}
+\npair{\left(
+        (\ms_S,\reg_S,\stk,\ms_\stk),
+        (\ms_T,\reg_T)\right)} \; \mid \; \\
     \begin{array}{l}
@@
     \end{array}
+  \end{multlined}
 \right\}
 \]
+\end{definition}
+Definition~\ref{def:exec-conf-rel} essentially says that two executable configurations are related when they are made out of related components.
+That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$- and $\rdata$ registers are related by the $\lrexj$ relation.
+Further, the two memory and the two register files should be related.
+This means that the executable configuration only contains words that respect memory invariants.
 
 % TODO Include central lemmas?
+\end{jversion}
 

### 2019-06-17 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Proof read the proving full abstraction section

@@
 Say, you manage to write something that can break memory invariants, then it would not be possible to read it back again as write permission, generally speaking, does not entail read permission.
-Iff the capability had read permission, then $\readCond{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants.
+If the capability had read permission, then $\readCond{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants.
 After all, if such a capability was obtained, memory invariants could be broken.
@@
 It should always be possible to write safe values, so we impose this as a lower bound.
-We do, however, impose a bound on what should be permitted to be written and that is a lower bound because it should at least be allowed to write anything that is safe.
-
-Before we get to the definition of $\writeCond{}$, we need to introduce one auxiliary definition.
-When a write capability makes a change to memory, the change must be permitted by the regions that govern the memory it has authority over for it to not break the invariant and thus be safe.
-Given a write capability, it should be possible to transform the memory from one allowed by the region to another also allowed by the region.
-However, only one memory address can be updated at a time, so in order to make such a transformation one has to update one address at a time.
-This should not break any memory invariants, so we require all the memories that we could go through to make the transformation to also be permitted by the region.
-This is captured by address stratification.
+
+A safe write capability must respect the memory invariant of the region that governs the memory the capability gives access to.
+Now consider the case, where the invariant permits two memory segments that differs in two or more addresses.
+In this case, the write capability cannot be used to transform the memory from one memory segment to the other because only one memory address can be updated at a time.
+If an adversary had such a capability, then it should be possible for them to transform the memory in a way that is consistent with the region.
+In other words, the adversarial code should be able to transform the memory segment to any memory segment permitted by the region.
+This is captured by address stratification (Definition~\ref{def:address-stratified}) which basically says that if a region permits two memory segments, then all the intermediate memory segments you may end up with when transforming one memory segment to the other must be permitted as will.
 \begin{definition}
@@
 \end{definition}
-We define $\writeCond{}$ as follows
+With address stratification defined, we define the write condition.
+\begin{definition}
 \[
@@
 \]
+\end{definition}
 The definition of $\writeCond{}$ is very similar to $\readCond{}$.
@@
 The two new condition are essentially the same as the $\readCond{}$ and $\writeCond{}$ except that they use the free stack sub-world and assume that the capability is linear as all stack capabilities are linear.
-Note that we do not have any condition that talks about the stack-frames sub world because we should never have a capability that allows us to directly read or write from that part of memory.
+Note that we do not have any condition that talks about the stack-frames sub world because we should never have a capability that allows us to directly read from or write to that part of memory.
+\begin{definition}
 \[
@@
 \]
+\end{definition}
+\begin{definition}
 \[
@@
 \]
+\end{definition}
 The final permission, we define conditions for is the execute permission.
@@
 We later define the $\lre$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
-An executable capability can have its range of authority shrunk or its current address changed which changes what it does and thus potentially whether it keeps memory invariants.
-For this reason, the condition requires that any capability with a derived range of authority and a current address that range is safe to use for execution.
-The $\execCond{}$ is quantified over all future worlds of the $\purePart{}$ of $W$. The $\purePart{}$ function strips the world of all spatial ownership.
-This means that the capability cannot depend on linear capabilities, and when we define the logical relation, we even require the executable capability to not be linear.
-It is not clear whether a linear executable capability would even be useful because there is no way to move it out of the $\pcreg$-register without crashing the execution.
-This may sound like an ideal option for constructing something that can be executed once, however, most programs rely on loading other capabilities or seal sets for the program counter which is not possible.
+An executable capability can have its range of authority shrunk or its current address changed which changes what instructions are executed and thus potentially whether the code respects memory invariants.
+For this reason, the condition requires that any executable capability with a derived range of authority and a current address in that range is safe to use for execution.
+The $\execCond{}$ is quantified over all future worlds of the $\purePart{}$ of $W$.
+We do not know when the executable capability will be used, so it should be safe even in the future when the memory has changed.
+The $\purePart{}$ function strips the world of all spatial ownership.
+This means that the capability cannot depend on linear capabilities and thus the contents of the stack.
+When we define the logical relation, we even require the executable capability to not be linear.
+It is not clear whether a linear executable capability would even be useful because it cannot be moved from the $\pcreg$-register without crashing the execution.
+This may sound like an ideal primitive for constructing something that can be executed once, however, most programs rely on loading other capabilities or seal sets using the program counter capability which is not possible when the program counter is linear.
 \begin{align*}
@@
       \forall n' < n, W' \future \purePart{W}\ldotp \forall b',e'\ldotp \forall \aaddr \in [b',e'] \subseteq A \ldotp\\
-      \quad \npair[n']{\stpair[.]{((\rx,\normal),\baddr',\eaddr',\aaddr)}{((\rx,\normal),\baddr',\eaddr',\aaddr)}} \in \lre(W')
+      \quad \npair[n']{\left( \arraycolsep=0pt\array{l}((\rx,\normal),\baddr',\eaddr',\aaddr),\\((\rx,\normal),\baddr',\eaddr',\aaddr)\endarray\right)} \in \lre(W')
     \end{array}
@@
 \end{align*}
-The $\readCond{}$ condition by itself allows many different regions and thus potentially many different memories.
-However, when we have a read capability with execute permission, we know that the capability must point to a piece of this code memory.
+The $\readCond{}$ condition by itself allows many different regions and thus potentially many different memory segments.
+However, when we have a read capability with execute permission, we know that the capability must point to a piece of code memory.
 For this reason, we define the $\xReadCond{}$ to capture the additional assumptions that we can make when a capability is executable.
@@
 It is always safe to posses data as data is just integers that grant no authority.
-Further, unlike capabilities, it is always possible to construct a new integer with the move instruction.
-
-The next case is capabilities that do not have a special representation on \srccm{} unlike stack pointers and return pointers.
-For two capabilities to be related, they should be syntactically equal. That is, they should have same range of authority, linearity and so one.
-Generally speaking, untrusted components should not have direct access to a trusted components code.
-For this reason, we require that capabilities must have a range of authority outside the trusted address space if they are to be related.
+Further unlike capabilities, it is always possible to construct a new integer with the move instruction.
+
+Next we have capabilities that do not have a special representation on \srccm{}, i.e.\ all capabilities but stack pointers and return pointers.
+For two capabilities to be related, they should be syntactically equal.
+That is, they should have the same range of authority, linearity and so one.
+Generally speaking, untrusted components should not have direct access to a trusted components code, so we require that capabilities must have a range of authority outside the trusted address space if they are to be related.
 The safety of a capability also depends on the world and whether the capability can be used to break the memory invariants of the world.
-For instance, if a capability has read-permission, then it should only by able to read safe values with it.
+For instance, if a capability has read-permission, then it should not be possible to read something unsafe, i.e.\ something that can break memory invariants.
 This condition and conditions for the other permissions are captured by the permission based conditions, so we use them to express the necessary conditions.
-That is, if a capability has read permission, then $\readCond{}$ must be satisfied, if it has writer permission, then the $\writeCond{}$ must be satisfied, and if it has execute permission, then $\execCond{}$ and $\xReadCond{}$ must be satisfied.
+That is, if a capability has read permission, then $\readCond{}$ must be satisfied, if it has write permission, then $\writeCond{}$ must be satisfied, and if it has execute permission, then $\execCond{}$ and $\xReadCond{}$ must be satisfied.
 If the capability has execute permission, then it must also be a non-linear capability.
-Finally, the capability cannot have read/write/execute permission because that would break the write-XOR-execute assumption.
+Finally, the capability cannot have read/write/execute permission because that would break the write-XOR-execute assumption, i.e.\ the code memory in non-writable and data memory is non-executable.
 
@@
 The $\xjmp{}{}$ takes a pair of sealed capabilities, so the authority of a sealed capability depends on what other sealed capabilities it might be used with.
+The seal invariant defines what may be sealed with a given seal and thus which sealables we should consider.
 As discussed, closure seals and return seals must be used in specific ways, so we have defined a seal invariant that dictates what may be sealed with a given seal.
-In order for a pair of \srccm{} and \trgcm{} seals to be in the unsafe value relation, they must be sealed with the same seal $\sigma$, and they must be related by the seal invariant $\sigma$.
+In order for a pair of \srccm{} and \trgcm{} seals to be in the untrusted value relation, they must be sealed with the same seal $\sigma$, and they must be related by the seal invariant $\sigma$.
 % Sealed capability only useful with pairing. Sealed caps must be permitted.
-Further, if we take any sealed capabilities related under $\seal$ in the seal invariant, then they should produce a safe result.
+Further, if we take any sealed capabilities related under $\sigma$ in the seal invariant, then they should produce a safe result.
 We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrexj$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
 % Linear and non-linear case. Affects the future world we consider.
-It should be safe to use the sealed capability both now and later.
-This means that it should be safe in the world $W$, but also in future worlds $W' \future W$.
+It should be safe to use the sealed capability both now and in the future.
+This means that it should be safe in the current world $W$, but also in future worlds $W' \future W$.
 However, this only goes for non-linear words.
@@
 For sets of seals to be related related in the untrusted relation they must be syntactically equal.
-Further, the seals in the set should be disjoint from the return seals and trusted closure seals ($\gsigrets$ and $\gsigcloss$ because trusted code relies on having the sole access to them.
+Further, the seals in the set should be disjoint from the return seals and trusted closure seals ($\gsigrets$ and $\gsigcloss$) because the trusted code relies on having the sole access to them.
 We do not know what an adversary may seal or what seal they may use, so, for every seal in the seal set, we require the seal invariant to be essentially equal to the untrusted value relation.
@@
 Two capabilities are related as trusted code pointers if they are non-linear, has a permission derivable from read-execute, and has a range of authority within the trusted address space, $\ta$.
-Further, we need to know that they are actually capability which is why we require the $\xReadCond{}$ to be satisfied.
+Further, we need to know that the capabilities actually point to a piece of code which is why we require the $\xReadCond{}$ to be satisfied.
 This makes sure that the region that governs the memory the capability points to is a code region.
 Note that even though the capability has read permission, we do not require the read condition to hold.
-The code memory contains trusted closure seals and return seals that we cannot let untrusted code have, so the read condition would not hold.
-It is, however, fine to let trusted code have access to such seals because we expect it to treat the seals reasonably.
+The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrv[]$, so the read condition would not hold.
+It is, however, fine to let trusted code have access to such seals because we expect the trusted code to treat the seals reasonably.
 \begin{figure}
@@
 Intuitively, two register files are related when they only contain safe words, i.e.\ words from the value relation.
-This raises the question ``which value relation''?
+This raises the question ``which value relation?''
 We only use the register file relation to relate register files for components we do not trust, so the answer is the untrusted value relation.
-The definition of the value relation is straightforward.
-It distributes the authority of the world among the registers and require each of the registers to contain a safe word with respect to the authority it is given.
+The definition of the register-file relation is straightforward.
+It distributes the authority of the world among the registers and requires each of the registers to contain a safe word with respect to the authority it is given.
 The register file never takes into account the $\pcreg$ and it can leave out further registers.
-We use this to not relate register content that we are going to overwrite with new values.
+We use this to not relate register content that will be overwritten anyway.
 We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$.
-That is, if we do not need to exclude extra registers, then we simply omit that argument.
-The register file relation is defined in \figurename~\ref{fig:reg-file-relation}.
+That is, if we do not need to exclude additional registers, then we simply omit that argument.
+The register file relation is defined in \figurename~\ref{fig:reg-file-rel}.
 
@@
 % TODO write section
-The expression relation $\lre$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e. construct configurations in the observation relation.
-The $\lre$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control-flow control-flow during execution when a $\tjmp{}$ instruction is executed.
+The expression relation $\lre$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e.
+the capabilities can be used to construct configurations in the observation relation.
+The $\lre$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control flow during execution when a $\tjmp{}$ instruction is executed.
 In the setting of \srccm{} and \trgcm{}, we can also use sealed capabilities to change the control flow by using the $\txjmp{}{}$ instruction.
-The $\txjmp{}{}$ updates the $\pcreg$ register and the $\rdata$ register, however, the $\lre$ relation only updates the $\pc$-register, so we cannot use $\lre$ to reason about sealed capabilities.
+The $\txjmp{}{}$ instruction updates the $\pcreg$ register and the $\rdata$ register, however, the $\lre$ relation only updates the $\pc$-register, so we cannot use $\lre$ to reason about sealed capabilities.
 Instead, we define the relation $\lrexj$ which relates two pairs of capabilities when they are safe to use with the $\txjmp{}{}$ instruction.
 
-In this case, related executions means that the observable effect of the executions should be permissible.
+Executions are related when the observable effect of the executions are permissible.
 The permissible observations are defined by the observation relation, so we define the expression relation in terms of the observation relation.
@@
 We lift the capabilities to configurations simply by plugging the two capabilities into the $\pcreg$-register of two configurations.
-We cannot pick arbitrary configurations because an arbitrary configuration may contain words that can be used to break memory invariants and create observable effects that we do not permit.
+We cannot pick arbitrary configurations because an arbitrary configuration may contain words that can be used to break memory invariants and thus create unacceptable observable effects.
 Instead, we need to pick configurations made out of related components, i.e.\ related register files and related memories that respect linearity.
@@
 It takes related memories and register files (ignoring the $\rdata$ register) and combines them into two configurations.
-Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lrexj$ relation.
+Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lro$ relation.
 The $\lrexj$ relation is defined in \figurename~\ref{fig:expr-rels}.
@@
 In order to show full abstraction (Theorem~\ref{thm:full-abstraction}), we need not only to relate the words on \srccm{} with words on \trgcm{} we also need to relate \srccm{} components with \trgcm{} components.
-Specifically, we say that two components are related when they are syntactically equal after all a \srccm{} component is in some sense the same as a \trgcm{} as we only see the difference during execution when a call happens and when we lift a component to a configuration which means that we need to introduce a stack pointer.
+Specifically, we say that two components are related when they are syntactically equal, after all, a \srccm{} component is in some sense the same as a \trgcm{} as we only see the difference during execution when a call happens and when we lift a component to a configuration where we need to introduce a stack pointer.
 However, we cannot take arbitrary components as they could potentially break memory invariants.
-For related components, we require that if the imports are satisfied with related words, then the resulting memory should be safe.
+For related base components, we require that if the imports are satisfied with related words, then the resulting memory should be safe.
 Further, related components must have safe exports.
-This roughly describes when a base components are related.
-A component with a main is related when the base component is related and the mains capabilities are in the exports.
+Components with a main are related when the base components are related and the main capabilities are in the public interface, that is they must be in the exports.
 
@@
 The full abstraction theorem (Theorem~\ref{thm:full-abstraction}) is stated in terms of contextual equivalence.
-Contextual equivalence (Definition~\ref{def:contextual-equivalence}) plugs two components with a context and requires equitermination of the resulting executable configurations.
-This means that we need to lift relatedness one step further, namely to the level of execution configurations.
+Contextual equivalence (Definition~\ref{def:contextual-equivalence}) plugs two components into a context and requires equitermination of the resulting executable configurations.
+This means that we need to lift relatedness one step further than the components, namely to the level of execution configurations.
 To this end, we define $\lrec$.
@@
 Definition~\ref{def:exec-conf-rel} essentially says that two executable configurations are related when they are made out of related components.
-That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$- and $\rdata$ registers are related by the $\lrexj$ relation.
-Further, the two memory and the two register files should be related.
+That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$ and $\rdata$ registers are related by the $\lrexj$ relation.
+Further, the two memories and the two register files should be related.
 This means that the executable configuration only contains words that respect memory invariants.
@@
 Using Lemma~\ref{thm:ftlr}, we can now proceed to proving Theorem~\ref{thm:full-abstraction} (full abstraction).
-First, we extend the logical relation into an omitted relation on components $(\var{comp}_S,\var{comp}_T) \in \lrcomp(W)$.
+
 % Essentially, it relates a component to itself if instantiating their imports with related values produces related exports and code memory satisfying the appropriate code region in the world.
@@
 \begin{proof}[Proof of Theorem~\ref{thm:full-abstraction}]
-  The proofs of both directions are similar, so we only show the right direction.
+  Both directions of the proof are similar, so we only show the right direction.
   To show the \trgcm{} contextual equivalence, assume w.l.o.g\ a well-formed context $\trg{\context}$ such that ${\plug{\trg{\context}}{\src{\comp_1}} \term[]{}}$.

### 2019-06-17 popl-2019-journal-v/discussion.tex
[POPLjv] Move discussion section to new file

--- 2019-05-07 popl-2019-journal-v/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,screen]{acmart}\settopmatter{}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986.10002989</concept_id>
-<concept_desc>Security and privacy~Formal security models</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10003006</concept_id>
-<concept_desc>Security and privacy~Systems security</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10010124.10010138</concept_id>
-<concept_desc>Theory of computation~Program reasoning</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Security and privacy~Formal security models}
-\ccsdesc[300]{Security and privacy~Systems security}
-\ccsdesc[300]{Theory of computation~Program reasoning}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-% \acmJournal{PACMPL}
-% \acmVolume{1}
-% \acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
-% \acmArticle{1}
-% \acmYear{2018}
-% \acmMonth{1}
-% \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-% \startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-% \setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-\usepackage[utf8]{inputenc}
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-% %% Some recommended packages.
-% \usepackage{booktabs}   %% For formal tables:
-%                         %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-\usepackage{wrapfig}
-
-\usepackage{placeins}
-
-\usepackage[disable]{todonotes}
-
-\input{preamble-illu}
-
-\input{../fully-abstract-comp/preamble}
-% Paper specific redefinitions of commands
-\renewcommand{\npair}[2][n]{#2}
-\renewcommand{\MemSeg}{\shareddom{MemSeg}}
-\renewcommand{\Reg}{\shareddom{RegFile}}
-\renewcommand{\RegName}{\shareddom{RegName}}
-\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
-\renewcommand{\encInstr}[1]{\plainfun{encode}{#1}}
-\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
-\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
-\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
-\renewcommand{\perm}{\var{p}}
-\renewcommand{\SealableCaps}{\shareddom{Sealables}}
-\renewcommand{\Cap}{\shareddom{NatTok}}
-\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
-\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
-\renewcommand{\comp}{\var{comp}}
-\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
-\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlr{F}(#3)}
-\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
-\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
-
-\renewcommand{\lre}[1][]{\typesetlr{E}^{#1}}
-\renewcommand{\lretrusted}[1][]{\lreg[#1]{\trusted}}
-\renewcommand{\lrexj}[1][]{\typesetlr{E}^{#1}_{\mathrm{xjmp}}}
-\renewcommand{\lrk}[1][]{\typesetlr{K}^{#1}}
-\renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
-\renewcommand{\lrr}[1][]{\lrrg[#1]{\untrusted}}
-\renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
-\renewcommand{\lro}[1][]{\typesetlr{O}^{#1}}
-\renewcommand{\lrol}{\lro[\preceq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
-\renewcommand{\lror}{\lro[\succeq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
-\renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
-\renewcommand{\lrv}[1][]{\lrvg[#1]{\untrusted}}
-\renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
-\renewcommand{\lrp}{\typesetlr{P}}
-
-\renewcommand{\lrrs}{\typesetlr{R}}
-\renewcommand{\lrm}{\typesetlr{M}}
-\renewcommand{\lrcomp}[1][]{\typesetlr{C}^{#1}}
-\renewcommand{\lrec}[1][]{\typesetlr{EC}^{#1}}
-\renewcommand{\lrheap}[1][]{\typesetlr{H}^{#1}}
-
-\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
-\renewcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
-\renewcommand{\stareg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta},#2}_{#1}}
-\renewcommand{\codereg}[2][\mathrm{code}]{\iota^{#1}_{#2}}
-\renewcommand{\staureg}[2][\stpair{\ms}{\ms}]{\iota^{\mathrm{sta,\lrv},#2}_{#1}}
-
-\newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
-\newcommand{\srcxjmpres}[1]{\plainfun{\srcalt{xjmpRes}}{#1}}
-\newcommand{\wdjud}[2][ ]{#1 \vdash #2}
-
-\newcommand{\trgcm}{\textsc{LCM}}
-\newcommand{\srccm}{\textsc{oLCM}}
-\newcommand{\extend}[1]{}
-\newcommand{\Rel}[1]{\mathrm{Rel}(#1)}
-\newcommand{\fers}[1][n]{\left(\nequal[#1]\right)_{#1=0}^{\infty}}
-\newcommand{\cofe}[2][n]{\left(#2,\fers[#1] \right)}
-\newcommand{\seq}[2][n]{\left(#2_{#1}\right)_{#1=0}^{\infty}}
-\setlength{\belowcaptionskip}{-5pt}
-
-\definecolor{Blue}{RGB}{92,26,143}
-\definecolor{OliveGreen}{RGB}{60,128,49}
-\newenvironment{jversion}%{}{}
-    {\color{OliveGreen}}{}
-\newenvironment{jversionsug}
-    {\begin{comment}}{\end{comment}}
-
-\newcommand{\ijversion}[1]%{#1}
-    {{\color{OliveGreen} #1}}
-\newcommand{\ijversionsug}[1]%{}
-    {{\color{Blue} #1}}
-
-%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
-%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
-%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
-%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
-%%%
-%%% The following is specific to POPL '19 and the paper
-%%% 'StkTokens: Enforcing Well-Bracketed Control Flow and Stack Encapsulation using Linear Capabilities'
-%%% by Lau Skorstengaard, Dominique Devriese, and Lars Birkedal.
-%%%
-\setcopyright{rightsretained}
-\copyrightyear{2019}
-\acmJournal{PACMPL}
-\acmYear{2019}
-\acmVolume{3}
-\acmNumber{POPL}
-\acmArticle{19}
-\acmMonth{1}
-\acmPrice{}
-\acmDOI{10.1145/3290332}
-
-\begin{document}
-
-%% Title information
-\title[\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
-% \titlenote{with title note}
-% \subtitle{Fully abstract overlay semantics}
-% \subtitlenote{with subtitle note}
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-\author{Lau Skorstengaard}
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \institution{Aarhus~University}
-  \country{Denmark}                    %% \country is recommended
-}
-\email{lask@cs.au.dk}
-
-%% Author with two affiliations and emails.
-\author{Dominique Devriese}
-\orcid{0000-0002-3862-6856}             %% \orcid is optional
-\affiliation{
-  \institution{Vrije~Universiteit~Brussel}           %% \institution is required
-  \country{Belgium}                   %% \country is recommended
-}
-\email{dominique.devriese@vub.be}         %% \email is recommended
-
-\author{Lars Birkedal}
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \institution{Aarhus~University}
-  \country{Denmark}                    %% \country is recommended
-}
-\email{birkedal@cs.au.dk}
-
-\begin{abstract}
-  We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
-  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
-  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalize and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
-\end{abstract}
-
-% %% 2012 ACM Computing Classification System (CSS) concepts
-% %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-% \begin{CCSXML}
-% <ccs2012>
-% <concept>
-% <concept_id>10011007.10011006.10011008</concept_id>
-% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-% <concept_significance>500</concept_significance>
-% </concept>
-% <concept>
-% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-% <concept_significance>300</concept_significance>
-% </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Software and its engineering~General programming languages}
-% \ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, fully abstract overlay semantics}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-\todo[inline]{TODOS ENABLED!!!}
-\todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be built.
-Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
-Generally, properties that hold in the source language but not in the target language need to be somehow enforced by the compiler.
-Two properties that hold in many high-level source languages are well-bracketed control flow and encapsulation of local state, but they are not enforced after compilation to assembly.
-
-Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
-At the assembly level, this property does not hold immediately. 
-Invoked functions get direct access to return pointers that they are supposed to jump to a single time at the end of their execution.
-There is, however, no guarantee that untrusted assembly code respects this intended usage.
-In particular, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.
-
-Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not be read or modified until the invoked function returns.
-At the assembly level, this property also does not hold immediately.
-The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
-However, untrusted assembly code is free to ignore this requirement and read or overwrite the local state of other stack frames.
-
-To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
-The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
-More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
-These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
-Capabilities come in different flavours.
-Memory capabilities allow reading from and writing to a block of memory.
-Additionally, capability machines offer some form of {\itshape object capabilities} that represent low-level encapsulated closures, i.e.
-a piece of code coupled with private state that it gains access to upon invocation.
-The concrete mechanics of object capabilities vary between different capability machines.
-% KJAA: This 'sealed' and 'common seal' business is unclear to me.
-For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
-Each of the two capabilities are sealed with a common seal which make them opaque.
-The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.
-
-To enforce WBCF and LSE on a capability machine, there are essentially two approaches.
-The first is to use separate stacks for mutually distrusting components, and a central, trusted stack manager that mediates cross-component invocations.
-This idea has been applied in CheriBSD (an operating system built on CHERI)~\citep{watson_capability_2015}, but it is not without downsides.
-First, it requires reserving separate stack space for all components, which scales poorly to large amounts of components.
-Also, in the presence of higher-order values (e.g., function pointers, objects), the stack manager needs to be able to decide which component a higher-order value belongs to in order to provide it the right stack pointer upon invocation.
-It is not clear how it can do this efficiently in the presence of large amounts of components.
-Finally, this approach does not allow passing stack references between components.
-
-A more scalable approach retains a single stack shared between components.
-Enforcing WBCF and LSE in this approach requires a way to temporarily provide stack and return capabilities to an untrusted component and to revoke them after it returns.
-While capability revocation is expensive in general, some capability machines offer restricted forms of revocation that can be implemented efficiently.
-For example, CHERI offers a form of {\itshape local} capabilities that can only be stored in registers or on the stack but not in other parts of memory.
-\citet{skorstengaard_reasoning_2017} have demonstrated that by making the stack and return pointer local, and by introducing a number of security checks and measures, the two properties can be guaranteed.
-In fact, a similar system was envisioned in earlier work on CHERI~\citep{watson2012cheri}. 
-% KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.
-
-In this work, we propose and study \stktokens{}: an alternative calling convention that enforces WBCF and LSE with a single shared stack.
-Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
-Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson2018ISA}.
-% KJAA: Perhaps make it clear by who and when/where?
-The hardware prevents these capabilities from being duplicated.
-We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to give them back on return.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
-
-A second contribution of this work is the way in which we formulate these two properties.
-Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
-Existing formulations are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
-We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
-It starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To this end, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
-This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
-As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
-To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
-More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
-
-Compared to~\citet{skorstengaard_reasoning_2017} that prove LSE and WBCF for a handful of examples,
-this approach expresses what it means to enforce the desirable properties in a general way and makes it clear that we can support a very general class of programs.
-Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
-The idea is that such a compiler could be proven fully abstract with respect to the well-behaved semantics of the target language, so that the proof could rely on native well-bracketedness and local stack frame encapsulation.
-Such an independent result could then be composed with ours to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.
-
-%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
-In this paper, we make the following contributions:
-% Presents the first formal model of a capability machine with linear capabilities (Section 2)
-% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
-% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
-\begin{itemize}
-\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
-\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
-\item We present a new way to formalize these guarantees based on a novel
-  technique called \textit{fully-abstract overlay semantics} and we prove LSE
-  and WBCF claims. This includes:
-  \begin{itemize}
-  \item \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
-  \item proving full-abstraction for the embedding of \srccm{} into \trgcm{} (Section~\ref{sec:fa-proof}) by
-  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
-  \end{itemize}
-\end{itemize}
-This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.
-
-%\paragraph{Outline} Blabla
-
-\section{A Capability Machine with Sealing and Linear Capabilities}
-\label{sec:cap-mach-w-seal-and-lin}
-% Brief intro to section and this cap machine:
-% Cap machine inspired by CHERI, but with linear capabilities
-In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
-For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
-% \lau{Do we have a good reason to assume an infinite address space for this work?
-%   We do not model malloc, so we do not need to worry about what happens when malloc runs out of memory.}
-
-%%% Capabilities
-The concept of a capability is the cornerstone of any capability machine.
-In its simplest form, a capability is a permission and a range of authority.
-The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
-% Permissions and Range of authority
-The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
-Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
-The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$) ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
-% Pointer and linearity
-In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
-The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
-% What is linear capabilities, split/splice
-A linear capability is a capability that cannot be duplicated.
-This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
-The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.
-
-\begin{wrapfigure}{r}{0.30\linewidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (rwx) {$\rwx$};
-    \node[main node] (rx) [below right of=rwx] {$\rx$};
-
-    \node[main node] (rw) [below left of=rwx,] {$\rw$};
-    \node[main node] (r) [below right of=rw] {$\readonly$};
-    \node[main node] (0) [below of=r] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-
-    (rw) edge (r)
-    (r) edge (0)
-
-    (rwx) edge (rx)
-
-    (rw) edge (rwx)
-    (r) edge (rx);
-  \end{tikzpicture}
-
-  \caption{Permission hierarchy}
-  \label{fig:perm-hier}
-\end{wrapfigure}%
-%
-%%% Seals
-
-Any reasonable capability machine needs a way to set up boundaries between security domains with different authorities.
-It also must have a way to cross these boundaries such that (1) the security domain we move from can encapsulate and later regain its authority and (2) the security domain we move to regains all of its authority.
-%The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015,watson_fast_2016}.
-% seals
-A sealed capability $\sealed{\sigma,\vsc}$ is a pair of a seal $\sigma$ and a capability $\vsc$.
-A sealed capability makes an underlying capability opaque which means that the underlying capability cannot be changed or used for the operations it normally gives permission to.
-In other words, the authority the underlying capability represents is encapsulated under the seal.
-In order to seal a sealable with a seal $\sigma$, it is necessary to have the authority to do so.
-The permission to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
-A set of seals is a capability that represents the authority to seal sealables with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In spirit of memory capabilities, a set of seals has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
-% On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
-As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}, an operation that operates on a pair of capabilities sealed with the same seal.
-The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part of the pair) available to the invoked code.
-The combination of sealed capabilities and \texttt{xjmp} gives (1) and (2).
-\begin{figure}[tb]
-  \centering
-  \[
-  \arraycolsep=1.4pt
-  \begin{array}{rrcl l rrcl}
-   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
-    \aendbnf \in & & & \Addr \cup \{\infty \} & & \sealendbnf \in & & & \Seal \cup \{\infty \}\\
-    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
-    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
-    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
-    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
-    \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
-\ms \in  &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
-    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
-  \end{array}
-\]
-\[
-  \arraycolsep=1.4pt
-\begin{array}{rcl}
-\multicolumn{3}{l}{    \arraycolsep=0pt
-      \com{r} \in  \tRegName \hspace{2.5cm}   \com{\rn} \defbnf \com{r} \mid \nats
-}\\
-  \Instr &\defbnf & \tjmp{\com{r}} \mid \tjnz{\com{r}}{\com{\rn}} \mid \tmove{\com{r}}{\com{\rn}} \mid \tload{\com{r}}{\com{r}} \mid \tstore{\com{r}}{\com{r}} \mid \tplus{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tminus{\com{r}}{\com{\rn}}{\com{\rn}} \mid\\
-         & & \tlt{\com{r}}{\com{\rn}}{\com{\rn}} \mid \tisptr{\com{r}}{\com{r}} \mid\tgetp{\com{r}}{\com{r}} \mid \tgetlin{\com{r}}{\com{r}} \mid \tgetb{\com{r}}{\com{r}} \mid \tgete{\com{r}}{\com{r}} \mid \tgeta{\com{r}}{\com{r}}  \mid \\
-  & & \tcca{\com{r}}{\com{n\rn}} \mid \tsetatob{\com{r}} \mid \trestrict{\com{r}}{\com{\rn}} \mid \tcseal{\com{r}}{\com{r}} \mid \txjmp{\com{r}}{\com{r}} \mid  \tsplit{\com{r}}{\com{r}}{\com{r}}{\com{\rn}} \mid\\
-      & & \tsplice{\com{r}}{\com{r}}{\com{r}} \mid \tfail \mid \thalt 
-\end{array}
-\]
-  \caption{The syntax of our capability machine with seals and linear capabilities.}
-  \label{fig:target-syntax}
-\end{figure}
-
-%%% Domains
-Words on \trgcm{} are capabilities and data (represented by integers $\ints$).
-We assume a finite set of register names $\RegName$ containing at least the registers $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$.
-We define register files as functions from register names to words.
-Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ partial functions).
-$\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
-An executable configuration is a register file and memory pair.
-
-\begin{figure}[p]
-  \centering
-  \begin{mathpar}
-    \inference{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \\
-        \baddr \le \aaddr \le \eaddr \and \perm \in \{\rwx,\rx\}
- }{ \Phi \step \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) }
-              \and
-              \inference{ \forall \Phi' \neq \failed \ldotp \Phi \nrightarrow \Phi'}
-                        {\Phi \step \failed }
-  \end{mathpar}
-  \[
-    \begin{array}{l}
-  %   \Phi \step
-  % \begin{cases}
-  %   \sem{\decInstr{\Phi.\mem(\aaddr)}}(\Phi) &
-  %   \arraycolsep=0pt
-  %     \begin{array}[t]{l}
-  %       \text{if }\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr) \wedge
-  %       \baddr \le \aaddr \le \eaddr \wedge \perm \in \{\rwx,\rx\}
-  %     \end{array} \\
-  %     \failed & \totherwise
-  % \end{cases}\\
-  \updPcAddr{\Phi} =
-  \begin{cases}
-    \Phi\updReg{\pcreg}{w} & \Phi(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \wedge w = ((\perm,\lin),\baddr,\eaddr,\aaddr+1)\\
-    \Phi  & \totherwise
-  \end{cases}\\
-  \linCons{w} =
-  \begin{cases}
-    0 & \isLinear{w} \\
-    w & \totherwise
-  \end{cases}\\
-  \xjmpres{c_1,c_2,\Phi} =
-  \begin{cases}
-    \Phi\updReg{\pcreg}{c_1}\updReg{\rdata}{c_2} & \nonExec{c_2} \\
-    \failed & \totherwise
-  \end{cases}
-  \end{array}
-\]
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.7cm}|}
-    \hline
-    i \in \Instr                                 & \sem{i}(\Phi) & Conditions\\
-    \hline
-    \halt                                        & \halted & \\
-    \hline
-    \fail                                        & \failed & \\
-    % \hline
-    % \move{r}{\rn}                                & \updPcAddr{\Phi\updReg{r}{\rn}} & $\rn \in \ints$\\
-    \hline
-    \move{r}{\rn}                                & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\Phi(\rn)}$ \\
-    \hline
-    \load{r_1}{r_2}                              & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{w_1}\\ &\update{\mem.\aaddr}{w_\aaddr})\endarray & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$ and $w_a = \linCons{w_1}$\\
-    \hline
-    \store{r_1}{r_2}                             & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ & \update{\mem.\aaddr}{\Phi(r_2)})\endarray & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
-    \hline
-    \geta{r_1}{r_2}                              & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\_,\aaddr)$ or $\Phi(r_2) = \seal{\_,\_,\aaddr}$, then $w = \aaddr$ and otherwise $w = -1$\\
-    \hline
-    \cca{r}{\rn}                                 &\updPcAddr{\Phi\updReg{r}{w}} & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
-    \hline
-    \jmp{r}    &\Phi\arraycolsep=0pt\array[t]{l}\updReg{r}{w}\\\updReg{\pcreg}{\Phi(r)}\endarray & $w = \linCons{\Phi(r)}$\\
-    \hline
-    \xjmp{r_1}{r_2}                              & \Phi' & $\Phi(r_1) = \sealed{\sigma,c_1}$ and $\Phi(r_2) = \sealed{\sigma,c_2}$ and $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$ and $\Phi' = \xjmpres{c_1,c_2,\Phi\updReg{r_1,r_2}{w_1,w_2}}$  \\
-    \hline
-    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_3)}$\\
-    \hline
-    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
-    \hline
-    \cseal{r_1}{r_2}                             & \updPcAddr{\Phi\updReg{r_1}{\vsc}} & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
-    \hline
-    \multicolumn{3}{|c|}{\dots} \\
-    \hline
-    \_                                           & \failed & \totherwise \\
-    \hline
-  \end{tabular}
-\caption{An excerpt of the operational semantics of \trgcm{}}
-  \label{fig:target-op-sem}
-\end{figure}
-% syntax (instructions)
-\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machines as well as capability-related instructions.
-The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
-The simplest of the capability instructions inspect the properties of capabilities: type (\texttt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
-The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
-The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
-Generally speaking, a capability machine needs an instruction for reducing the range of authority of a capability.
-Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two rather than reducing the range of authority (\texttt{split}).
-The reverse is possible using \texttt{splice}.
-Sealables can be sealed using \texttt{cseal} and pairs of sealed capabilities can be unsealed by crossing security boundaries (\texttt{xjmp}, see below).
-Finally, \trgcm{} has instructions to signal whether an execution was successful or not (\texttt{halt} and \texttt{fail}).
-
-% opsem
-The operational semantics of \trgcm{} is displayed in Figure~\ref{fig:target-op-sem}.
-The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
-The executed instruction is determined by the capability in the $\pcreg$ register, i.e.\ $\Phi(\pcreg)$ (we write $\Phi(r)$ to mean $\Phi.\reg(r)$).
-In order for the machine to take a step, the capability in the $\pcreg$ must have a permission that allows execution, and the current address of the capability must be within the capability's range of authority.
-If both conditions are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
-The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
-In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
-The instructions that stop execution or change the flow of execution do not use $\updPcAddr{}$.
-For instance, the \texttt{halt} and \texttt{fail} instructions are simply interpreted as the $\halted$ and $\failed$ configurations, respectively, and they do not use $\updPcAddr{}$.
-
-% move w/ lin
-The \texttt{move} instruction simply moves a word from one register to another.
-It is, however, complicated slightly by the presence of the non-duplicable linear capabilities.
-When a linear capability is moved, the source register must be cleared to prevent duplication of the capability.
-This is done uniformly in the semantics using the function $\linCons{}$ that returns $0$ for linear capabilities and is the identity for all other words.
-When a word $w$ is transferred on the machine, then the source of $w$ is overwritten with $\linCons{w}$ which clears the source if $w$ was linear and leaves it unchanged otherwise.
-In the case of \texttt{move}, the source register $\rn$ is overwritten with $\linCons{\Phi(\rn)}$.
-
-% load w/ lin (and store)
-The \texttt{store} and \texttt{load} instructions are fairly standard.
-They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority.
-Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
-In this case, we require that the memory capability used for loading also has write-permission.
-
-% geta, cca
-The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
-\texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
-Note that this instruction does not need to use $\linCons{}$ like the previous ones, because it modifies the capability in-place, i.e.\ the source register is also the target register.
-The \texttt{jmp} instruction is a simple jump that just sets register $\pcreg$.
-
-%% Sealing
-The operational side of the sealing in \trgcm{} consists of two instructions: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
-% cseal
-Given a sealable $\vsc$ and a set of seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
-% xjmp
-Apart from dealing with linearity, \texttt{xjmp} takes a pair of sealed capabilities, unseals them, and puts one in the $\pcreg$ register and the other in the $\rdata$ register, but only if they are sealed with the same seal and the data capability (the one placed in $\rdata$) is non-executable.
-A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
-Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
-This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
-Opaque sealed capabilities encapsulate a security domain's local state and authority, and they only become accessible again when control is transferred to the security domain.
-Some care should be taken for sealing because reusing the same seal for multiple closures makes it possible to jump to the code of one closure with the environment of another.
-% xjmp to unseal
-\trgcm{} does not have an instruction for unsealing capabilities directly, but it can be (partially) simulated using \texttt{xjmp}.
-
-% split/splice
-Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
-For normal capabilities, reduction of authority can be done without actually giving up any authority by duplicating the capability first.
-With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
-In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
-The \texttt{split} instruction takes a capability with range of authority $[\var{base},\var{end}]$ and an address $n$ and creates two new capabilities, with $[\var{base},n]$ and $[n+1,\var{end}]$ as ranges of authority.
-Everything else, i.e.\ permission, linearity and current address, is copied without change to the new capabilities.
-With \texttt{split}, we can reduce the range of authority of a linear capability without losing any authority as we retain it in the second capability.
-The \texttt{splice} instruction essentially does the inverse of \texttt{split}.
-Given two capabilities with adjacent ranges of authority and the same permissions and linearity, \texttt{splice} splices them together into one capability.
-The two instructions work in the same way for seal sets.
-We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
-This would also allow linear capabilities to have aliases, but only by linear capabilities with disjoint permissions.
-
-\begin{jversion}
-% The remaining instructions
-The interpretation of the remainder of the instructions are displayed in Appendix~\ref{app:instr-interpretation}.
-The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all project information about capabilities.
-The $\tgetb{}{}$ instruction projects the base address of the range of authority, and $\tgete{}{}$ projects the end address of the range of authority.
-The linearity of a permission is projected with $\tgetlin{}{}$, and finally the permission is projected with $\tgetp{}{}$.
-The instructions $\tgetb{}{}$ and $\tgete{}{}$ also works on sets of seals.
-The instruction $\tisptr{}{}$ gives the type of a word which allows programs to be defensive in the sense that they can check whether a word is a capability before they use the word as one.
-\trgcm{} also has arithmetic instructions $\tplus{}{}{}$, $\tminus{}{}{}$, and $\tlt{}{}{}$.
-The latter instruction compares two numbers and writes 1 or 0 to a target register depending on whether one number is less than the other.
-The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set of seals) to the base address of the range of authority (or range of seals).
-This instruction makes it possible to work relatively to the base address of a capability rather than with fixed addresses.
-While this instruction is not strictly necessary as it can be emulated with other instruction, it still seems useful in the case where you receive capabilities from an untrusted source.
-Finally, we have the $\trestrict{}{}$ instruction which  restricts the permission of a capability but only according to the $\le$ relation.
-\end{jversion}
-
-\begin{jversion}
-  \subsection{The purpose of sealing}
-  % Sealing example:
-  To motivate the necessity of an encapsulation mechanism like sealing, consider the following example.
-  We want to transfer control to code that we do not trust, and we want to give them the means to return to us.
-  That is, we need to give them a return capability.
-  If we did not have an encapsulation mechanism, our only option would be to give them an executable capability for the address we want them to return to.
-  The code we give the return capability could use it as intended, but it could also manipulate the return capability to point to a different point and return to an unintended point of our code.
-  Further in order to be able to retrieve our capabilities from before transferring control, we would have to store them somewhere accessible from the return capability.
-  However, the untrusted code would have access to all this through the return capability because it does not give less authority to the untrusted code than it does to us.
-  For this reason, any reasonable capability machine must have an encapsulation mechanism to enable programs to make boundaries between security domains.
-
-  In the example, if we use sealing, we can seal the return capability which does two things.
-  It prevents the untrusted code from changing the target of the return capability forcing them to return to the point we intended.
-  It also prevents the untrusted code from reading our capabilities.
-  All in all, this means that we can transfer control to another domain without giving up our capabilities or handing them over.
-
-  \trgcm{} does not have a direct unsealing instruction, but it is still possible to do some amount of unsealing.
-  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains the $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
-  Now assume part of the code would like to have access to $w$ in $\rdata$.
-  Take a capability for this code, possibly by adjusting the pc, and seal it with $\sigma$.
-  Now you have the data part and the code part of a sealed capability pair, so you can $\txjmp{}{}$ to it which unseals the data capability and puts it in $\rdata$ for your code to use.
-  Sealed executable capabilities cannot be unsealed in the same way because $\txjmp{}{}$ fails if the data capability is executable.
-  However, given a sealed executable capability $\sealed{\sigma,c}$ and a set of seals that contains $\sigma$, we can still construct the data part of the sealed capability pair.
-  This means that we can execute the code the capability points to in the presence of any data we can construct.
-
-  Sealing is meant for encapsulation, but it gives us very weak encapsulation properties if seals are not private as the unsealing example illustrates.
-  For this reason, it is important that the system is setup, so each component has access to seals that are unique to that component.
-  We return to this in \sectionname~\ref{sec:well-form-reas}.
-\end{jversion}
-
-\begin{jversion}
-\subsection{Decoding and encoding functions}
-The operational semantics of the capability machine uses $\decInstr{}$ to decode instructions.
-We also assume an $\encInstr{}$ which is used to put programs onto the capability machine.
-Rather than specifying a decode function and an encode function, we assume that they are given with certain properties.
-The $\decInstr{} : \Word \fun \Instr$ should be injective and surjective for all non-$\fail$ instructions.
-Further, it should decode all capabilities as the $\fail$ instruction, i.e.\
-\[
-\forall c \in \Cap \ldotp \decInstr{c} = \fail
-\]
-The $\encInstr{} : \Instr \fun \ints$ function should be injective, and it should be defined so $\decInstr{}$ is its left inverse, i.e.
-\[
-\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
-\]
-
-These assumptions are sufficient to construct program examples and run them on the machine.
-The $\encInstr{}$ function allows us to specify the program ``abstractly'' in terms of instructions rather than machine words.
-As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
-%This is not to say that encoding is an easy problem. There are many things to get.
-
-The machine also assumes decode and encode functions for permissions $\decPerm{} : \Perm \fun \ints$ and $\encPerm{} : \ints \fun \Perm$.
-We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$ and surjective. For $\encPerm{}$, we assume it is surjective and that it does not encode anything to the error value -1, i.e.
-\[
-  \forall \perm \in \Perm \ldotp \encPerm{\perm} \neq -1
-\]
-For linearity we assume similar functions.
-
-Finally, the machine uses an encode function for word types $\encType{}$ in the interpretation of $\tisptr{}{}$.
-This function encodes each kind type of word as an integer.
-This is very much like the previous functions.
-It encodes each kind of word differently and all words of the same kind to the same integer.
-In Section~\ref{subsec:overlay-semantics}, we present an overlay semantics which changes the set of words compared to this machine.
-The overlay semantics also has a type encoding function which must encode the new words, so they correspond to their representation on this machine.
-For instance, the encoding of the type of a stack-pointer in the overlay semantics should be the same as the encoding of a capability on this machine.
-All the new words correspond to a word on this machine, and their encodings should match.
-\end{jversion}
-
-\subsection{Components, linking, programs, and contexts }
-\label{subsed:components-linking}
-The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
-To clarify this, we introduce notions of components and programs from which we construct executable configurations.
-A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
-It has exports that can satisfy the imports of other components.
-% Structure
-A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses (addresses governed by a linear capability).
-The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
-The exports are words each associated with a symbol.
-%base = library
-A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
-The latter can be seen as a program that still needs to be linked with libraries.
-% Linking .
-Components are combined into new components by linking them together, as long as only one is an incomplete program with a main.
-Two components can be linked when their memories, seals, and linear addresses are disjoint.
-They are combined by taking the union of each of their constituents.
-For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
-The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
-
-% Programs
-We can now define the notion of a program as well as a context.
-\begin{definition}[Programs and Contexts]
-  \label{def:program-and-context}
-  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
-  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
-\end{definition}
-% Getting an executable from a component - postpone this to later, but mention:
-% write xor exec
-How a program is initialised to create an executable configuration, is discussed in Section~\ref{sec:form-secur-with}.
-% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
-%Further, we will only consider a specific class of well-formed components, but we will get back to this.
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for details.
-\begin{figure}[tb]
-  \centering
-  \[
-    \begin{array}{rl c rl c rl}
-    s &\in \Symbol & \phantom{make} &
-    \mathrm{import} &\mathrel{::=} a \mapsfrom s & \phantom{make} &
-    \mathrm{export} &\mathrel{::=} s \mapsto w\\
-    \comp_0 & \multicolumn{7}{l}{\mathrel{::=} (\mscode,\msdata,\overline{\mathrm{import}},\overline{\mathrm{export}},\sigrets,\sigcloss,A_\linear)}\\
-    \comp & \multicolumn{7}{l}{\mathrel{::=} \comp_0 \mid  (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})}
-  \end{array}
-\]
-\vspace{0.5cm}
-\begin{mathpar}
-  \inference{
-    \comp_0 = (\mscode[1], \msdata[1], \overline{\var{import}_1}, \overline{\var{export}_1}, \sigrets[1], \sigcloss[1],A_{\linear,1})\\
-    \comp_0' = (\mscode[2], \msdata[2], \overline{\var{import}_2}, \overline{\var{export}_2}, \sigrets[2], \sigcloss[2],A_{\linear,2})\\
-    \comp_0'' = (\mscode[3], \msdata[3], \overline{\var{import}_3}, \overline{\var{export}_3}, \sigrets[3], \sigcloss[3],A_{\linear,3})\\
-    \mscode[3] = \mscode[1] \uplus \mscode[2] \\
-    \msdata[3] = (\msdata[1] \uplus \msdata[2])[a \mapsto w \mid (a \mapsfrom s) \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}), (s \mapsto w) \in \overline{\var{export}}_3] \\
-    \overline{\var{export}_3} = \overline{\var{export}_1} \cup \overline{\var{export}_2}&
-    \overline{\var{import}_3} = \{ a \mapsfrom s \in (\overline{\var{import}_1} \cup \overline{\var{import}_2}) \mid s \mapsto \_ \not\in \overline{\var{export}_3} \}\\
-    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
-    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2] &
-    A_{\linear,3} = A_{\linear,1} \uplus A_{\linear,2}\\
-    \dom(\mscode[3]) \mathrel{\#} \dom(\msdata[3]) & \sigrets[3] \mathrel{\#} \sigcloss[3]
-  } {
-    \comp_0'' = \comp_0 \bowtie \comp_0'
-  }
-  \and
-  \inference{
-    \comp_0'' = \comp_0 \bowtie \comp_0'
-  }{
-    (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \comp_0 \bowtie (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
-  }
-  \end{mathpar}
-% \begin{definition}[Plugging a program into a context]
-%   When $\var{comp'}$ is a context for component $\comp$ and $\comp' \bowtie \comp \rightsquigarrow \Phi$, 
-%   then we write $\plug{\comp'}{\comp}$ for the execution configuration $\Phi$.
-% \end{definition}
-  \caption{Components and linking of components.}
-  \label{fig:target-component-and-linking}
-\end{figure}
-
-%\FloatBarrier
-% \begin{itemize}
-% \item present our \emph{target language} and its operational semantics (excerpts)
-% \item mention roughly what components look like
-% \end{itemize}
-\section{Linear Stack and Return Capabilities}
-\label{sec:stktokens-explained}
-In this section, we introduce our calling convention \stktokens{} that ensures LSE and WBCF.
-We will gradually explain each of the security measures \stktokens{} takes and motivate them with the attacks they prevent.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-
-      % Stack pointer 1
-      \begin{scope}
-        \clip (4.6,4) rectangle (9,13);
-        \capbracebot{(4.6,4)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 1}
-      \end{scope}
-      Stack pointer 2
-      \begin{scope}
-        \clip (4.8,4) rectangle (9,13);
-        \capbrace{(5.2,9)}{(5.2,13.5)}{adv. stack\\\footnotesize cap. 2}
-      \end{scope}
-
-      \draw[red,thick,->] (4.5,10) to[out=0,in=50] node[midway,right] {} (7.5,7.3);
-      \draw[red,thick,->] (7,7.3) to[out=115,in=0] node[midway,right] {} (4.5,8);
-    \end{tikzpicture}
-    \caption{An adversary uses a previous stack frame's stack pointer.}
-    \label{fig:stack-ptr-abuse}
-  \end{subfigure}
-  \begin{subfigure}{0.18\linewidth}
-    \phantom{testtestes}
-  \end{subfigure}
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-      \fill[red,draw=red,opacity=.4] (-.1,3.9) rectangle (4.6,9.1);
-      \draw[red,thick,->] (4.5,10.5) to[out=0,in=0] node[midway,right,align=center] {return\\ to wrong\\ return ptr} (4.5,3);
-    \end{tikzpicture}
-    \caption{An adversary jumps to a previous stack frame's stack pointer.}
-    \label{fig:ret-ptr-abuse}
-  \end{subfigure}
-
-  \caption{Possible ways to abuse stack and return capabilities.}
-  \label{fig:stack-ret-ptr-abuse}
-\end{figure}
-
-\stktokens{} is based on a traditional single stack, shared between all components.
-To explain the technique, let us first consider how we might already add extra protection to stack and return pointers on a capability machine.
-First, we replace stack pointers with stack capabilities.
-When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
-Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
-% KJAA: Again with the sealed pair. Again, this idea is unclear.
-They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
-
-%informally explain how an adversary may try to abuse stack and return caps
-While the above adds extra protection, it is not sufficient to enforce WBCF and LSE.
-Untrusted callees receive a stack capability and a return pair that they are supposed to use for the call.
-However, a malicious callee (which we will refer to as an adversary\footnote{See Section~\ref{sec:well-form-reas} for more details on our attacker model.}) can store the provided capabilities on the heap in order to use them later.
-Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
-In both examples our component and an adversarial component have been taking turns calling each other, so the stack now contains four stack frames alternating between ours and theirs.
-The figure on the left (Figure~\ref{fig:stack-ptr-abuse}) illustrates how we try to ensure LSE by restricting the stack capability to the unused part before every call to the adversary.
-However, restricting the stack capability does not help when we, in the first call, give access to the part of the stack where our second stack frame will reside as nothing prevents the adversary from duplicating and storing the stack pointer.
-Generally speaking, we have no reason to ever trust a stack capability received from an untrusted component as that stack capability may have been duplicated and stored for later use.
-In the figure on the right (Figure~\ref{fig:ret-ptr-abuse}), we have given the adversary two pairs of sealed return capabilities, one in each of the two calls to the adversarial component.
-The adversary stores the pair of sealed return capabilities from the first call in order to use it in the second call where they are not allowed.
-The figure illustrates how the adversarial code uses the return pair from the first call to return from the second call and thus break WBCF.
-
-% Informally explain how we prevent this using linear capabilities
-As the examples illustrate, this naive use of standard memory and object capabilities does not provide sufficient guarantees to enforce LSE and WBCF.
-The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
-Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver no longer has it or a copy of it as it is non-duplicable.} that can be used to prevent this from happening.
-
-% Stack capability linear
-The linear capabilities are put to use by requiring the stack capability to be linear.
-On call, the caller splits the stack capability in two: one capability for their local stack frame and another one for the unused part of the stack.
-The local stack frame capability is sealed and used as the data part of the sealed return pair.
-The capability for the remainder of the stack is given to the callee.
-% Prevention of left attack non-aliasing of linear capabilities - token like (ensuring LSE)
-Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
-This means that an adversary would need the stack capability produced by the caller in order to access their local data.
-The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
-This is illustrated in Figure~\ref{fig:stack-ptr-abuse-prev} and prevents the issue illustrated in Figure~\ref{fig:stack-ptr-abuse}.
-
-% Prevention of attack 2
-In a traditional calling convention with a single stack, the stack serves as a call stack keeping track of the order calls where made in and thus in which order they should be returned to.
-A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
-However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
-This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
-In the presence of adversarial code, we need some mechanism to enforce that the order of the call stack is kept.
-One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
-The callee would have to provide this token on return to prove that it is allowed to return to the caller, and on return the token would be taken back by the caller to prevent it from being spent multiple times.
-As it turns out, the stack capability for the unused part of the stack can be used as such a token in the following way:
-On return the callee has to give back the stack capability they were given on invocation.
-When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack or not.
-They do this by attempting to restore the stack capability from before the call by splicing the return token with the stack capability for the local stack frame which at this point has been unsealed again.
-If the splice is successful, then the caller knows that the two capabilities are adjacent.
-On the other hand, if the splice fails, then they are alerted to the fact that their stack frame may not be the topmost.
-\stktokens{} uses this approach; and as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}, it prevents the issue in Figure~\ref{fig:ret-ptr-abuse} as the adversary does not return a spendable token when they return.
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
-
-      % Stack pointer 1
-      \sealsymb{(4.9,5.5)}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{adv. stack\\\footnotesize cap. 1}
-      % return cap
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{data return\\\footnotesize cap. 2}
-     %  Stack pointer 2
-      \begin{scope}
-        \clip (4.8,4) rectangle (9,13);
-        \lincapbrace{(4.6,9)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 2}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{The non-duplicable linear stack capability for the trusted code's
-      stack frame and the opacity of sealed capabilities ensures LSE.}
-    \label{fig:stack-ptr-abuse-prev}
-  \end{subfigure}
-  \begin{subfigure}{0.18\linewidth}
-    \phantom{testtestes}
-  \end{subfigure}
-  \begin{subfigure}{0.4\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
-      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}
-
-      \lincapbrace{(4.6,2)}{(4.6,4)}{data return\\\footnotesize cap. 1}
-
-     %  Stack pointer 2
-      \begin{scope}
-        \clip (4.8,4) rectangle (9,13);
-        \lincapbrace{(4.6,9)}{(4.6,13.5)}{adv. stack\\\footnotesize cap. 2}
-      \end{scope}
-
-      \redcross{(5,4)}
-    \end{tikzpicture}
-    \caption{The trusted caller fails to splice the stack capability returned by
-    the adversary with the capability for the trusted caller's local stack frame.}
-    \label{fig:ret-ptr-abuse-prev}
-  \end{subfigure}
-  \caption{Abuse of stack and return capabilities prevention.}
-\end{figure}
-
-% Non-empty trusted stack frames.
-In order for a call to have a presence on the call stack, its stack frame must be non-empty.
-We cannot allow empty stack frames on the call stack, because then it would be impossible to tell whether the topmost non-empty stack frame has an empty stack frame on top of it.
-Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in \stktokens{} these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
-This means that a caller using \stktokens{} needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
-There is also a more practical reason for a \stktokens{} caller to make sure their stack frame is non-empty: They need a bit of the stack capability in order to perform the splice that verifies the validity of the return token.
-
-% known stack base
-% + more
-At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
-However, this still does not ensure that the caller's stack frame is on top of the call stack.
-The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
-An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
-This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
-The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
-The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
-At this point (Figure~\ref{fig:stack-base-abuse-b}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
-
-For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
-In \stktokens{}, the base address of the stack capability is fixed as a compile-time constant (Note: the stack grows downwards, so the base address of the stack capability is the top-most address of the stack).
-The caller verifies the validity of the return token by checking whether the base address of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
-In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
-
-% Where do we get the stack pointer from, how can we know it is linear
-% The other attack.
-% At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
-% However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
-% In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
-% To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
-% The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
-% The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
-% Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
-% At this point, the adversary has everything they need to return from either of the two callbacks.
-% Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
-% From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
-% They are both correct as the adversarial component has created two call stacks, one for each of the components.
-% In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
-% We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
-% A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
-% This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.
-
-In \stktokens{}, the stack memory is only referenced by a single linear stack capability at the start of execution.
-Because of this, the return token can be verified simply by checking its base address and splicing it with the caller's stack frame.
-There is no need to check linearity because only linear capabilities to this memory exist.
-
-% Return pointers/return seals
-The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
-Both of the capabilities in the pair are sealed with the same seal.
-% One seal per return point
-All call points need to be associated with a unique seal (a return seal) that is only used for the return capabilities for that particular call point.
-The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that gave rise to the stack frame, breaking WBCF.
-For similar reasons, we cannot allow return seals to be used to seal closures.
-% Don't be stupid
-Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
-This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a capability for reading, either directly or indirectly, a return seal from memory).
-
-% Note about them vs us, we do nothing that they couldn't do
-We have sometimes phrased the description of the \stktokens{} calling scheme in terms of ``them vs us''.
-This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
-However, \stktokens{} is a modular calling scheme: no restriction is put on adversarial components that we do not expect trusted components to meet.
-Specifically, we are going to assume that both trusted and adversarial components are initially syntactically well-formed (described in more detail in Section~\ref{sec:well-form-reas}) which basically just restrict adversarial components to not break machine guarantees initially (e.g.\ no aliases for linear capabilities or access to seals of other components).
-This means that any component can ensure WBCF and LSE by employing \stktokens{}.
-
-% KJAA: ^-- Assuming that there is a global stack address capability, no?
-% Can I "suddenly" begin to impose \stktokens{}?
-
-% Summary of CC
-To summarise, \stktokens{} consists of the following measures:
-\begin{enumerate}
-\item \label{item:check-stkb} Check the base address of the stack capability before and after calls.
-\item \label{item:non-empty-sf} Make sure that local stack frames are non-empty.
-\item \label{item:return-data} Create token and data return capability on call: split the stack capability in two to get a stack capability for your local stack frame and a stack capability for the unused part of the stack.
-The former is sealed and used for the data part of the return pair.
-\item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, every return seal is only used for one call site, and they are not leaked.
-\end{enumerate}
-Item~\ref{item:check-stkb}-\ref{item:return-code} are captured by the code in Figure~\ref{fig:call-code} , except for checking stack base before calls.
-We do not include this check because it only needs to happen once between two calls, so that the check after a call suffices if the stack base is not changed subsequently.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.23\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \actadv{(0,4)}{(4.5,7)}{}
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-      \lincapbracebot{(4.6,4)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-a}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
-      \node[opacity=0.5,rotate=-45] at (2.25,10) {\footnotesize Sent to trusted};
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{}
-
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-      \lincapbracebot{(4.6,7)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-b}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \actadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
-      \node[opacity=0.5,rotate=-45] at (2.25,11) {\footnotesize Returned to adv.};
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{}
-
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{}
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-c}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \actsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (4.6,13.1);
-        \draw[fill=gray!50] (0,9) rectangle (4.5,13.5);
-      \end{scope}
-
-      \lincapbrace{(4.6,2)}{(4.6,7)}{}
-
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{}
-
-      \fill[red,draw=red,opacity=.4] (-.1,6.9) rectangle (4.6,9.1);
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-d}
-  \end{subfigure}
-  \caption{Partial return token used to return out of order.}
-  \label{fig:stk-base-abuse}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-{\footnotesize
-\[
-  \begin{array}{r >{$}p{5.1cm}<{$} r l}
-\multicolumn{2}{l}{\text{// Ensure non-empty stack.}}            & \multicolumn{2}{l}{\text{// Clear tmp registers and jump.}}\\
-1 :& \tmove{\rtmp{1}}{42}                                        &14: & \tmove{\rtmp{1}}{0}\\
-2 :& \tstore{\rstk}{\rtmp{1}}                                    &15: & \txjmp{r_1}{r_2}\\
-3 :& \tcca{\rstk}{(-1)}                                          &\multicolumn{2}{l}{\text{// The following is the return code.}}\\
-\multicolumn{2}{l}{\text{// Split stack in local stack frame and unused.}}      &\multicolumn{2}{l}{\text{// Check that returned stack pointer has base $\stkb$.}}\\
-4 :& \tgeta{\rtmp{1}}{\rstk}                                     &16: & \tgetb{\rtmp{1}}{\rstk}\\
-5 :& \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}                     &17: & \tminus{\rtmp{1}}{\rtmp{1}}{\stkb}\\
-\multicolumn{2}{l}{\text{// Load the call seal.}}                &18: & \tmove{\rtmp{2}}{\pcreg}\\
-6 :& \tmove{\rtmp{1}}{\pcreg}                                    &19: & \tcca{\rtmp{2}}{5}\\
-7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
-8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \rtmp{2}{1}\\
-9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
-\multicolumn{2}{l}{\text{// Seal the local stack frame.}}        &23: & \tfail \\
-10:& \tcseal{\rretd}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Splice with capability for local stack frame.}}\\
-\multicolumn{2}{l}{\text{// Construct code return pointer.}}     &24: & \tsplice{\rstk}{\rstk}{\rdata} \\
-11:& \tmove{\rretc}{\pcreg}                                      &\multicolumn{2}{l}{\text{// Pop 42 from the stack}}\\  
-12:& \tcca{\rretc}{5}                                            &25: & \tcca{\rstk}{1}\\
-13:& \tcseal{\rretc}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Clear tmp register}}\\
-      &    &26: & \tmove{\rtmp{2}}{0}\\
-  \end{array}
-\]
-}
-\caption{
-  % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
-  % Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
-  % Instructions 1-3: ensure that the stack is non-empty.
-  % Instructions 4-5: split stack pointer in two (stack pointer for callee done).
-  % Instructions 6-9: Load the seal designated for this call.
-  % Instruction 10: Seal capability for private stack to create data return pointer.
-  % Instructions 11-13: Seal capability for return point to create code return pointer.
-  % Instructions 14-15: Clean up temporary register and jump.
-  % Instructions 16-26: The return code.
-  % Instructions 16-23: check the stack base.
-  % Instruction 24: join the stack pointer.
-  % Instructions 25-26: pop 42 from the stack.
-  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set of seals it uses and $\offsigma$ the offset in the set of seals to the call seal.
-  $\stkb$ is the globally agreed on stack base.
-  There are some magic numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
-  Line 7: $-5$, offset from line 6 (where $\pcreg$ was copied into $\rtmp{1}$) to line 1.
-  Line 12: $5$, offset to the return address.
-  Line 19: $5$, offset to fail.
-  Line 21: offset to address after fail.
-}
-  \label{fig:call-code}
-\end{figure}
-
-% \begin{itemize}
-% \item informally explain how an adversary may try to abuse stack and return caps
-% \item informally explain how we prevent this using linear capabilities
-% \item use the tikz pictures from the PriSC presentation to explain all of this
-% \end{itemize}
-
-\section{Formulating Security with a Fully Abstract Overlay Semantics}
-\label{sec:form-secur-with}
-As mentioned, the \stktokens{} calling convention guarantees well-bracketed control flow (WBCF) and local state encapsulation (LSE).
-However, before we can prove these properties, we need to know how to even formulate them.
-Although the properties are intuitively clear and sound precise, formalizing them is actually far from obvious.
-
-Ideally, we would like to define the properties in a way that is
-\begin{enumerate}
-\item {\itshape intuitive} \label{def-prop:intuitive}
-\item {\itshape useful for reasoning:} we should be able to use WBCF and LSE when reasoning about correctness and security of programs using \stktokens{}. \label{def-prop:useful}
-\item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
-\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
-\label{def-prop:complete}
-\item {\itshape potentially scalable}: although dynamic code generation and multi-threading are currently out of scope, the formalization should, at least potentially, extend to such settings.\label{def-prop:scalable}
-\end{enumerate}
-
-Previous formalisations in the literature are formulated in terms of a static control flow graph~\cite[e.g., ][]{Abadi2005Theory}.
-While these are intuitively appealing (\ref{def-prop:intuitive}), it is not clear how they can be used to reason about programs (\ref{def-prop:useful}) or other compiler passes (\ref{def-prop:reusable}), they lack temporal safety guarantees (\ref{def-prop:complete}) and do not scale (\ref{def-prop:scalable}) to settings with dynamic code generation (where a static control flow graph cannot be defined).
-\citet{skorstengaard_reasoning_2017} provide a logical relation that
-can be used to reason about programs using their calling convention
-(\ref{def-prop:useful},\ref{def-prop:reusable}), but it is not intuitive (\ref{def-prop:intuitive}), there is no argument for completeness (\ref{def-prop:complete}), and it is unclear whether it will scale to more complex features (\ref{def-prop:scalable}).
-
-We contribute a new way to formalise the properties using a novel approach we call fully abstract overlay semantics.
-The idea is to define a second operational semantics for programs in our target language.
-This second semantics uses a different abstract machine and different run-time values, but it executes in lock-step with the original semantics and there is a very close correspondence between the state of both machines.
-
-The main difference between the two semantics, is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
-Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
-For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.
-
-The fact that \stktokens{} enforces LSE and WBCF is then formulated as a theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
-The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
-Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than components which the trusted component interacts with in the well-behaved overlay semantics.
-In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
-More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics iff they are indistinguishable to arbitrary other components in the overlay semantics.
-
-Our formal results are complicated by the fact that they only hold on a sane initial configuration of the system and for well-behaved components that respect the basic rules of the calling convention.
-For example, the system should be set up such that seals used by components for constructing return pointers are not shared with other components.
-We envision distributing seals as a job for the linker, so this means our results depend on the linker to do this properly.
-As another example, a seal used to construct a return pointer can be reused but only to construct return pointers for the same return point.
-Different seals must be used for different return points.
-Such seals should also never be passed to other components.
-These requirements are easy to satisfy: components should request sufficient seals from the linker, use a different one for every place in the code where they make a call to another component, and make sure to clear them from registers before every call.
-The general pattern is that $\stktokens{}$ only protects components that do not shoot themselves in the foot by violating a few basic rules.
-In this section, we define a well-formedness judgement for the syntactic requirements on components as well as a reasonability condition that semantically disallows components to do certain unsafe things.
-Well-formedness is a requirement for all components (trusted and untrusted), but the reasonability requirement only applies to trusted components, i.e.\ those components for which we provide LSE and WBCF guarantees.
-
-\begin{figure}[b]
-  \centering
-  \[
-    \arraycolsep=1.4pt
-    \begin{array}{rcl}
-      \src{\SealableCaps} & \defbnf& \SealableCaps \mid \src{\stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}} \mid\\
-                          & &   \src{\retptrd(\basebnf,\aendbnf)} \mid \src{\retptrc(\basebnf,\aendbnf,\addrbnf)}\\
-      \multicolumn{3}{c}{
-      \begin{array}{lcrclcr}
-        \src{\StkFrame} & \defeq & \src{\Addr \times \MemSeg} & \phantom{skipskipsip} & \src{\Stack} & \defeq & \src{ \StkFrame^*}
-      \end{array}
-                                                                                                                }\\
-      \src{\ExecConf} & \defeq & \Mem \times \Reg \; \src{\times \; \Stack \times \MemSeg} \\
-    \end{array}
-  \] 
-\[
-  \begin{array}{rclcrcl}
-    \src{\Instr} & \defbnf &  \Instr \mid \scall{\offpc,\offsigma}{r}{r}&\phantom{skipskipskip}&
-    \offpc,\offsigma & \in & \nats
-  \end{array}
-\]
-\caption{The syntax of \srccm{}.
-  \srccm{} extends \trgcm{} by adding stack pointers, return pointers, and a built-in stack.
-  Everything specific to the overlay semantics is written in blue.
-}
-  \label{fig:source-syntax}
-\end{figure}
-
-% see e-mail discussion: we do not see a good way to implement this TODO.
-% \todo[inline]{Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.}
-
-\subsection{Overlay Semantics}
-\label{subsec:overlay-semantics}
-% Describe the things added to the syntax
-The overlay semantics \srccm{} for \trgcm{} views part of the memory as a built-in stack (Figure~\ref{fig:source-syntax}).
-% Built-in stack
-To this end, it adds a call stack and a free stack memory to the executable configurations of \trgcm{}.
-The call stack is a list with all the stack frames that are currently inaccessible because they belong to previous calls.
-Every stack frame contains encapsulated stack memory as well as the program point that execution is supposed to return to.
-The free stack memory is the active part of the stack that has not been claimed by a call and thus can be used at this point of time.
-% Stack pointer
-In order to distinguish capabilities for the stack from the capabilities for the rest of the memory, \srccm{} adds stack pointers.
-A stack pointer has a permission, range of authority, and current address, just like capabilities on \trgcm{}, but they are always linear.
-% Return pointers
-The final syntactic constructs added by \srccm{} are the code and data return pointers.
-The data return pointer corresponds to some stack pointer (which in turn corresponds to a linear capability), and the code return pointer corresponds to some capability with read-execute permission.
-Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to on the underlying machine.
-On \srccm{}, return pointers are generated by calls from the capabilities they correspond to on \trgcm{}, and they are turned back to the capabilities they correspond to upon return.
-% Even though the return pointers both correspond to capabilities with read permission, neither can be used for reading. 
-% Say, we allowed reads through a data return pointer, then that would correspond to reading from some stack memory that is encapsulated in a stack frame which would break LSE.
-
-The opaque nature of the return pointers is reflected in the interpretation of the instructions common to both \trgcm{} and \srccm{} as \srccm{} does not add special interpretation for them in non-\texttt{xjmp} instructions.
-Stack pointers, on the other hand, need to behave just like capabilities, so \srccm{} adds new cases for them in the semantics, e.g.\ \texttt{cca} can now also change the current address of a stack pointer as displayed in Figure~\ref{fig:source-op-sem}.
-Similarly, \texttt{load} and \texttt{store} work on the free part of the stack when provided with a stack pointer.
-A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics has on the underlying machine.
-In other words, there should only be stack pointers for the stack memory.
-
-% Back to high-level view
-As discussed earlier, our formal results only provide guarantees for components that respect the calling convention.
-Untrusted components are not assumed to do so.
-To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
-Only instructions at these addresses can be interpreted as the \srccm{} native call and push frames to the call stack which guarantees LSE and WBCF.
-The constant $\ta$ is a parameter of the \srccm{} step relation.
-Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
-
-Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
-This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
-The sequence of instructions have to correspond to a call, i.e.\ the
-instructions in Figure~\ref{fig:call-code} ({\footnotesize  $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is always $26$, i.e.\ the number of instructions).
-Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.
-
-% Call/return
-The interpretation of $\scall{\offpc,\offsigma}{r_1}{r_2}$ is also shown in Figure~\ref{fig:source-op-sem} and essentially does the following:
-The registers $r_1$ and $r_2$ are expected to contain a code-data pair sealed with the same seal and the unsealed values are invoked by placing them in the $\pcreg$ and $\rdata$ registers, respectively.
-The current active stack and the stack capability are split into the local stack frame of the caller and the rest.
-$\scall{}{}{}$ also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
-The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of sealed return capabilities.
-The return capabilities are sealed with the seal designated for the call.
-
-The return capabilities, $\retptrc$ and $\retptrd$ are sealed and can only be used using the \texttt{xjmp} instruction, to perform a return.
-When this happens, the topmost call stack frame $(\opc,\ms_\var{local})$ is popped from the call stack.
-In order for the return to succeed, the return address in the code return pointer must match $\opc$, and the range of addresses in the data return pointer must match the domain of the local stack.
-If the return succeeds, the stack pointer is reconstructed, and the local stack becomes part of the active stack again.
-
-\srccm{} supports tail calls.
-A tail call is a call from a caller that is done executing, and thus doesn't need to be returned to or preserve local state.
-This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in call.
-To perform a tail call, the caller simply transfers control to the callee using \texttt{xjmp}.
-The tail-callee should return to the caller's caller, so the caller leaves the return pair they received for the callee to use.
-
-It is important to observe that the operational semantics of \srccm{} natively guarantee WBCF (well-bracketed control flow) and (local stack encapsulation) for calls made by trusted components.
-By inspecting the operational semantics of \srccm{}, we can see that it never allow reads or writes to inactive stack frames on the call stack.
-The built-in call for trusted code pushes the local stack frame to the inactive part of the stack, together with the return address. 
-Such frames can be reactivated by \texttt{xjmp}ing to a return capability pair, but only for the topmost stack frame and if the return address corresponds to the one stored in the call stack.
-In other words, WBCF and LSE are natively enforced in this semantics.
-
-\begin{figure}[htb]
-  \centering
-  \begin{mathpar}
-    \inferrule{\src{\Phi(\pcreg) = ((\perm,\_),\baddr,\eaddr,\aaddr)} \\
-               \src{\lbrack \aaddr ,\aaddr + \calllen - 1\rbrack \subseteq \ta}  \\
-               \src{\lbrack\aaddr,\aaddr + \calllen-1\rbrack \subseteq [\baddr,\eaddr]} \\
-               \src{\perm \in \{\rwx,\rx\}} \\
-               \src{\Phi.\mem(\aaddr,\dots,\addr + \calllen-1) = \scall[0]{\offpc,\offsigma}{r_1}{r_2} \cdots \scall[\calllen-1]{\offpc,\offsigma}{r_1}{r_2}}
- }
-              { \src{\Phi}  \; \src{\step[\src{\ta,\stkb}] \sem{\scall{\offpc,\offsigma}{r_1}{r_2}}(\Phi)} }
-  \end{mathpar}
-  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.6cm}|}
-    \hline
-    i \in \src{\Instr}                                 & $\sem{i}(\Phi)$ & Conditions\\
-    \hline
-    \halt                                        & $\halted$ & \\
-    \hline
-    \multicolumn{3}{|c|}{\dots\text{ (the operational semantics of \trgcm)}} \\
-    \hline
-    \store{r_1}{r_2}                             & $\srcalt{\var{updPc}} \sourcecolor\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\update{\ms_\stk.\aaddr}{\Phi(r_2)})\endarray$  & \srcalt{$\Phi(r_1) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$ and $\aaddr \in \dom(\ms_\stk)$}\\
-    \hline
-    \cca{r}{\rn}                                 &$\srcalt{\updPcAddr{\Phi\updReg{r}{w}}}$ &  \srcalt{$\Phi(\rn) = n \in \ints$ and $\Phi(r) = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w = \stkptr{\perm,\baddr,\eaddr,\aaddr+n}$} \\
-    \hline
-    \arraycolsep=0pt\array{c}
-    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}}
-    \endarray &
-        $\srcalt{\var{xjmpRes}(c_1,c_2,}$
-        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1,r_2}{w_1,w_2}\\
-            &\updReg{\rretc}{s_c}\\
-            &\updReg{\rretd}{s_d}\\
-            &\updReg{\rstk}{c_\stk}\\
-            &\update{\ms_\stk}{\ms_{\stk,\var{rest}}}\\
-            &\update{\stk}{\stk'}\\
-            \endarray\right))}$
-       &
-      \srcalt{
-         $\begin{multlined}
-           \ms_{\stk,\var{local}},c_\var{local},\ms_{\stk,\var{rest}},c_\stk =\\ \var{splitStack}(\Phi.\reg(\rstk), \Phi.\ms_\stk) \text{ and }
-         \end{multlined}$
-    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
-    $\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
-    $\sigma =\begin{multlined}[t]
-      \var{getCallSeal}(\\
-      \Phi.\reg(\pcreg), \Phi.\mem,\offpc,\offsigma)  \text{ and }
-    \end{multlined}$
-    $s_c,s_d = \var{sealReturnPair}(\sigma,c_\opc,c_\var{local})$ and
-    $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
-                        $\Phi.\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
-                        }\\
-    \hline 
-    \multicolumn{3}{|c|}{\dots} \\
-    \hline
-    \_                                           & $\failed$ & \totherwise \\
-    \hline
-  \end{tabular}
-\begin{multline*}
-  \srcxjmpres{c_1,c_2,\Phi} = \\
-  \left\{ 
-    \arraycolsep=1.4pt
-    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
-      \arraycolsep=0pt
-    \begin{array}{rl}
-      \Phi & \updReg{\pcreg}{c_1}\\
-           & \updReg{\rdata}{c_2}
-    \end{array}
-    &\phantom{mak} & $\nonExec{c_2}$ \srcalt{and $c_1 \neq \retptrc(\_)$ and $c_2 \neq \retptrd(\_)$}\\
-      \arraycolsep=0pt
-      \array[c]{rl}
-      \Phi&\updReg{\pcreg}{c_\opc}\\
-          &\updReg{\rstk}{c_\stk}\\
-          &\updReg{\rdata}{0}\\
-          &\update{\stk}{\stk'}\\
-          &\update{\ms_\stk}{\ms_\stk \uplus \ms_\var{local}}
-      \endarray
-           & &
-               $\arraycolsep=0pt\array{l}
-        \src{(\opc, \ms_\var{local}) :: \stk'= \Phi.\stk \tand} \\
-      \src{c_1 = \retptrc(\baddr,\eaddr,\opc)}\\
-      \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk) \wedge \dom(\ms_\var{local}) = [\aaddr_\stk,\eaddr_\stk]}\\
-      \src{c_\stk = \var{reconstructStackPointer}(\Phi.\reg(\rstk),c_2) \tand}\\
-      \src{c_\opc =  ((\rx,\normal),\baddr,\eaddr,\opc)} 
-      \endarray$\\
-      \failed &  & \totherwise
-    \end{array} 
-\right.
-\end{multline*}
-  \caption{An excerpt of the operational semantics of \srccm{} (some details omitted). Auxiliary definitions are found in Figure~\ref{fig:source-op-sem-aux}. }
-  \label{fig:source-op-sem}
-\end{figure}
-
-\begin{figure}
-    \begin{multline*}
-      \src{\var{splitStack}(\stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}, \ms_\stk) = \ms_{\stk,\var{local}},c_\var{local\_data},\ms_{\stk,\var{unused}},c_\stk} \mathit{\ iff\ } \\
-      \left\{ 
-      \begin{array}{l}
-      \src{\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk }\\
-      \src{\ms_{\stk,\var{local}} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} }\\
-      \src{\ms_{\stk,\var{unused}} = \ms_\stk|_{[\baddr_\stk,\aaddr_\stk-1]} }\\
-      \src{c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} }\\
-      \src{c_\var{local\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) }\\
-      \end{array}
-      \right.
-    \end{multline*}
-    \begin{align*}
-      \src{ \var{setupOpc(((\_,\_),\baddr,\eaddr,\aaddr))}} &\ \src{= \opc, c_\opc} \mathit{\ iff\ }
-      \left\{ 
-      \begin{array}{l}
-      \src{\opc = \aaddr + \calllen \wedge{}}\\
-      \src{c_\opc = \retptrc(\baddr,\eaddr,\opc) \wedge{}}
-      \end{array}
-       \right.\\
-      \src{\var{getCallSeal}(c_\pcreg,\mem,\offpc,\offsigma)} &\ \src{= \sigma} \mathit{\ iff\ }
-      \left\{
-      \begin{array}{l} 
-      \src{c_\pcreg = ((\_,\_),\baddr,\eaddr,\aaddr) \wedge \baddr \leq  \aaddr+\offpc \leq \eaddr \wedge{}} \\
-      \src{\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr} \wedge{} \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \wedge{}}\\
-      \src{\sigma = \sigma_\aaddr + \offsigma }
-      \end{array}
-      \right.\\
-     \src{\var{sealReturnPair}(\sigma,c_\opc,c_\var{local})} &\ \src{= \sealed{\sigma,c_\opc},\sealed{\sigma,c_\var{local}}}
-    \end{align*}%
-  % \begin{multline*}
-  %   \src{\var{callConditions}(r_1,r_2,\offpc,\offsigma,(\mem,\reg,\stk,\ms_\stk)) = (\mem,\reg',\stk',\ms_{\stk,\var{rest}})}\mathit{\ iff}\\
-  %     \quad\quad\left\{ 
-  %       \begin{array}{l}
-  %         \src{\var{sealedPairArgument}(\reg,r_1,r_2,c_1,c_2)}\\
-  %     \src{\ms_{\stk,\var{local}},\ms_{\stk,\var{rest}},c_\stk,c_\var{local\_data} = \var{splitStack}(\reg, \ms_\stk) \wedge{}}\\
-  %     \src{\opc, c_\opc =\var{setupOpc(\reg)} \wedge{}}\\
-  %     \src{\stk' = (\opc,\ms_{\stk,\var{local}}) :: \stk \wedge{}}\\
-  %     \src{\sigma = \var{getCallSeal}(\reg,\mem,\offpc,\offsigma) \wedge{}}\\
-  %     \src{s_c,s_d = \var{sealReturnPair}(\sigma,c_\var{local\_data},c_\opc) \wedge{}}\\
-  %     \src{w_1,w_2 = \linCons{\reg(r_1),\reg(r_2)} \wedge{}}\\ 
-  %     \src{\reg' =\reg\updReg{r_1,r_2}{w_1,w_2}\updReg{\rretc,\rretd}{s_c,s_d}\updReg{\rstk,\rtmp{1}}{c_\stk,0}}
-  %     \end{array}\right.
-  % \end{multline*}
-  \begin{multline*}
-    \src{\var{reconstructStackPointer}(\stkptr{\rw, \stkb, \aaddr_\stk-1,\_},\retptrd(\aaddr_\stk,\eaddr_\stk)) =}\\ \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk}} \mathit{\ iff\ }
-        \src{\stkb \leq \addr_\stk }
-   \end{multline*}
-
-  \caption{Auxiliary definitions used in the operational semantics of \srccm{}.}
-  \label{fig:source-op-sem-aux}
-\end{figure}
-% \begin{figure}
-% % \[
-% %     \src{\stkptr{\rw,\stkb,\eaddr_\stk,\aaddr_\stk} = \var{tokenToStkPtr}(\retptrd(\aaddr_\stk,\eaddr_\stk)}
-% % \]
-% % \[
-% %     \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame} (\stk,c_1,c_2)}
-% %     \mathit{\ iff\ }
-% %       \left\{ \array{l}
-% %         \src{\stk = (\opc, \ms_\var{local\_stk}) :: \stk' \wedge{}} \\
-% %         \src{c_1 = \retptrc(\_,\_,\opc) \wedge{}} \\
-% %         \src{c_2 = \retptrd(\aaddr_\stk,\eaddr_\stk)\wedge{}} \\
-% %         \src{\dom(\ms_\var{local\_stk}) \subseteq [\aaddr_\stk,\eaddr_\stk]}
-% %         \endarray
-% %      \right.
-% % \]
-% % \[
-% %   \src{\var{retPtrForMatchLocalStack(\retptrd(\aaddr_\stk,\eaddr_\stk),\ms_\var{local})}} \mathit{\ iff\ } \src{\dom(\ms_\var{local} \subseteq [\aaddr_\stk,\eaddr_\stk])}
-% % \]
-
-% % \begin{multline*}
-% %     \src{(\mem,\reg',\stk',\ms_\stk') = \var{returnCondition}(c_1,c_2,(\mem,\reg,\stk,\ms_\stk))}\mathit{\ iff\ }\\
-% %       \quad\quad\left\{ 
-% %       \begin{array}{l}
-% %       \src{\stk', \ms_\var{local\_stk} = \var{popCallStackFrame}(\stk,c_1,c_2)}\\
-% %       \src{c_\stk = \var{reconstructStackPointer}(\reg,c_2)}\\
-% %       \src{c_\opc =  \var{tokenToPc}(c_1)} \\
-% %       \src{\ms_\stk' = \ms_\stk \uplus \ms_\var{local\_stk}} \\
-% %       \src{\reg' =\reg\updReg{\pcreg,\rstk}{c_\opc,c_\stk}\updReg{\rdata,\rtmp{1},\rtmp{2}}{0,0,0}}
-% %       \end{array}\right.
-% %   \end{multline*}
-
-%   \caption{The return condition in the operational semantics of \srccm{}.}
-%   \label{fig:source-op-sem-return}
-% \end{figure}
-\subsection{Well-Formed Components}
-\label{sec:well-form-reas}
-% Well-formed components
-% \begin{figure}[htb]
-%   \centering
-%   \begin{mathpar}
-%   \inference{
-%     \dom(\mscode) = [\baddr,\eaddr]&
-%     [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
-%     \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
-%     \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \mscode \\
-%     \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)} & \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
-%     A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-%     \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigrets,\sigcloss \vdash_{\mathrm{comp-value}} \msdata(a)\\
-%     \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} & \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}}& \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
-%     \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}\\
-%     \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} & (\dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset)&
-%     \dom(\msdata) \mathrel{\#} \ta
-%   }{
-%     \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
-%   }
-%   \and
-%   \inference{
-%     \comp_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
-%     \vdash \comp_0 & (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
-%   }{
-%     \vdash (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
-%   }
-% \end{mathpar}
-%   \caption{Exceprt of well-formedness judgement}
-%   \label{fig:well-formedness}
-% \end{figure}
-% Well-formedness highlights:
-% Main exported
-% Code memory:
-% - Only instructions and seals
-% - calls associated with return seals unique return seal
-% - only uses seals specified for components
-% - (No linear access)
-% Data memory:
-% - Data
-% - Capabilities fora data memory (linearity respected)
-% - Sealed caps sealed with clos seal
-% Exports
-% - Any Word valid in data memory or
-% - Sealed capability for code (sealed with clos seal)
-% Imports
-% - Address must be in data memory
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
-The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
-
-\begin{jversion}
-  Informally, a well-formed component respects Write-XOR-Execute.
-  This means that there are no execute capabilities for the data memory and no write capability for the code memory.
-  The data memory of a well-formed component respects linearity and it may come with preset data structures.
-  It may, however, not contain capabilities for the code memory and sets of seals.
-  All the available seals of a component should reside in the code memory.
-  Both adversarial and trusted components can be well-formed.
-  Adversarial components are components where the range of addresses for the code memory is disjoint from the trusted address space Further, adversarial code does not have access to return seals.
-  On the other hand, trusted components must make return seals available to all its calls in a way that is consistent with \stktokens{}.
-  This means that each return seal is only used for one call.
-  Finally, well-formed components can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
-\end{jversion}
-
-% We first highlight the main points of the $\wdjud[\ta]{\comp}$ judgement.
-% For components with a main pair, the main pair must come from the exports and the remainder of the component must be well-formed.
-% As a reminder, a base component looks like this: $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$.
-% The return seals $\sigrets$ are the seals supposed to be usedto seal return pointers and the closure seals $\sigcloss$ all the other seals in a component.
-% If a component is adversarial, i.e.\ the domain of the code memory $\mscode$ is disjoint from the set of trusted addresses $\ta$, then there should be no return seals.
-% The code memory may contain sets of seals but only for the seals in $\sigrets$ and $\sigcloss$.
-% Other than that, code memory only contains instructions in the form of integers.
-% When a sequence of instructions in the code memory corresponds to a call (Figure~\ref{fig:call-code}), the call must have access to the return seal it specifies.
-% The return seal also needs to be unique to that call; that is, no other call can specify the same seal as its return seal.
-% The data memory $\msdata$ may contain data, capabilities, and sealed capabilities.
-% The capabilities in the data memory can only have authority over the data memory itself.
-% This allows components to have initial data structures.
-% In order to respect Write-XOR-Execute, the capabilities in data memory cannot have execute permission.
-% The capabilities also need to respect linearity which means that linear capabilities cannot be aliased by any other capability, and they must be for a range of the predetermined linear addresses $A_\linear$.
-% The sealed capabilities must be sealed with a closure seal, and the sealable can be anything that is allowed to reside in data memory.
-% The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
-
-\begin{jversion}
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
-The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
-
-%TODO distinguish closure seals from return seals.
-The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
-The \textsc{Main} rule requires the component to have a main pair picked from the exports.
-Further, the component should be a base component that satisfies judgement.
-The \textsc{Base} judgement makes sure that the component has a bit of structure.
-Roughly speaking, the \textsc{Base} rule requires the following:
-\begin{itemize}
-\item \textit{Code and data memory are disjoint.}
-\item \textit{Code memory is enclosed in a memory segment $\ms_\var{pad}$ that contains zero's.} This prevents the memory of two code memories from being spliced together.
-\item \textit{Data memory can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.} The component judgement ensure that the linear address space of a component is at most referenced by one capability.
-This means that any capability for the linear address space is unique.
-\item \textit{There is a partition $A_\var{own}$ of $A_\var{linear}$ that assigns a set of linear-addressed to each address of data memory.} The partitions are used by the component word judgement ($\vdash_\mathrm{comp-word}$) which ensures linear capabilities only address the addresses in the linear address space that is owned by the address where the capability resides in memory.
-\item \textit{All import addresses are part of the data memory.} 
-\item \textit{Import and export symbols are disjoint.}
-\item \textit{One of the following two are true}
-  \begin{itemize}
-  \item \textit{The code memory is disjoint from the trusted address space $\ta$ and in that case there are no return seals.} In this case, the component contains untrusted code.
-We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals.
-  \item \textit{The code memory is part of the trusted address space $\ta$.} In this case, the component contains trusted code.
-We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
-  \end{itemize}
-\item \emph{The data memory is disjoint from the trusted address space $\ta$.} We are not looking to execute data, so the trusted addresses never include the data memory addresses.
-\item \textit{The code memory satisfies the component-code judgement (\figurename~\ref{fig:well-formed-code}).}
-\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).
-In particular, this must be with respect to the linear addresses assigned to this address.
-}
-\item \textit{All the exports satisfies the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
-\end{itemize}
-
-\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ the words that are well-formed in a component's data memory.
-The judgement is defined by three rules: \textsc{W-Data}, \textsc{W-Capability}, and \textsc{W-Sealed-Capability}.
-The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
-The \textsc{W-Capability} rule specifies that all well-formed capabilities in data memory must at most have read and write permission.
-Further, if the capability is linear, then the range of authority must be within the linear address space owned by this address.
-Finally, if it is non-linear, then it must be in the non-linear address space.
-The \textsc{W-Sealed-Capability} specifies that well-formed sealed capabilities in data memory are sealed with a closure seal and whatever is sealed is itself a well-formed word.
-Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
-However, a component is free to place sealed data capabilities in memory during execution.
-
-\figurename~\ref{fig:well-formed-export} defines the well-formed exports $\vdash_{\mathrm{comp-export}}$.
-\textsc{E-Word} says that any well-formed word for data memory is well-formed as an export.
-In particular, this means that sealed capabilities for the data memory are acceptable exports which allows components to export the data part of a sealed capability pair.
-The \textsc{E-Sealed-Code} rule allows the code part of a sealed capability pair to be export.
-In particular, it allows a capability with read-execute permission and its range of authority within the component's code address space to be exported when it has been sealed with a closure seal.
-Together, these two rules allow components to have closures that they export.
-
-\figurename~\ref{fig:well-formed-code} defines the well-formed code memories $\vdash_{\mathrm{comp-code}}$.
-The well-formed code memories are defined by two judgements.
-One judgement considers the entire memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ while the other considers the contents of each code memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
-Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each of its words in isolation.
-The well-formedness of code memory depends on whether seals for calls are located in the place we expect them to be located.
-The $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$ judgement has two rules.
-The \textsc{C-Seal} judgement requires address $a$ in $\mscode$ to contain a set of seals that is a subset of the closure seals and the return seals.
-\textsc{C-Instr} requires address $a$ to contain an integer, which may be interpreted as instruction.
-If address $a$ is the first address of a series of instructions that may be interpreted as a call, and all the addresses are within the trusted address space, then there must be a set of seals available at the address the call in question expects it and that set of seals must contain the return seal that corresponds to the return seal of this call.
-The return seal in question must be in the $\sigrets[\mathrm{owned}]$ to ensure that no other call uses it.
-In order to ensure that return seals are at most used by one call, the $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ judgement takes all the available return seals and partitions them over code addresses.
-This means that no other call can use the same seal.
-Further, return seals cannot be used to seal non-return capabilities, so the \textsc{C-Mem} judgement requires the set of return seals to be disjoint from the set of closure seals.
-To limit the amount of corner cases we have to consider, we require the code memory to have no \textit{hidden calls}.
-\lau{Do we actually use the hidden calls in the proofs?
-I vaguely recall this was added to make sure that splicing would not create a call.
-However, we also have the padding which, at least in part, takes care of this.}
-\begin{definition}[No hidden calls]
-  \label{def:no-hidden-calls}
-  We say that a memory segment $\mscode$ has no hidden calls iff
-  \[
-    \begin{array}{l}
-      \forall \aaddr \in \dom(\mscode) \ldotp \\
-      \quad \forall i \in [0,\calllen - 1] \\
-      \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
-      \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1] \wedge\\
-      \qquad \qquad \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
-      \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
-      \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
-    \end{array}
-  \]
-\end{definition}
-This definition says that if code memory contains an instruction that may be part of a call, then either it is a part of a call or there is some other instruction that witnesses the fact that it is not part of a call.
-Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set of seals available.
-
-\begin{figure}
-\begin{mathpar}
-  \inferrule*[Right=Base]{
-    \dom(\mscode) = [\baddr,\eaddr]\\
-    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
-    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
-    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}\\
-    \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
-    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} \\
-    \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}} \\
-    \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
-    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} \\
-    (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset) \\
-    \dom(\msdata) \mathrel{\#} \ta \\
-    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
-    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
-    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
-  }{
-    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
-  }
-  \and
-  \inferrule*[Right=Main]{
-    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
-    \ta \vdash \var{comp}_0 \\
-    (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
-  }{
-    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
-  }
-\end{mathpar}
-\caption{Well-formedness judgement.}
-\label{fig:well-formed}
-\end{figure}
-
-\begin{figure}[h]
-  \begin{subfigure}[h]{1.0\linewidth}
-    \begin{mathpar}
-  \inferrule*[Right=C-Seals]{
-    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \\ [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
-  }{
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
-  }
-  \and
-  \inferrule*[Right=C-Instr]{
-    \mscode(a) \in \ints\\
-{    \begin{multlined}
-      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
-      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
-    \end{multlined}
-}  }{
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
-  }
-  \and
-  \inferrule*[Right=C-Mem]{
-    \text{$\mscode$ has no hidden calls}\\
-    \sigrets \mathrel{\#} \sigcloss \\
-    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
-    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
-    \forall a \in \dom(\mscode) \ldotp 
-    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
-    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
-  }{
-    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
-  }
-    \end{mathpar}
-    \caption{Code well-formedness.}
-    \label{fig:well-formed-code}
-  \end{subfigure}
-
-  \begin{subfigure}[h]{1.0\linewidth}
-    \begin{mathpar}
-      \inferrule*[Right=W-Data]{
-        z \in \ints
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} z
-  }
-  \and
-  \inferrule*[Right=W-Capability]{
-    \permbnf \sqsubseteq \rw\\
-    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}} \\\\
-    \lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
-  }
-  \and
-  \inferrule*[Right=W-Sealed-Capability]{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \vsc \\
-    \sigma \in \sigcloss
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \sealed{\sigma,\vsc}
-  }
-    \end{mathpar}
-    \caption{Word well-formedness.}
-    \label{fig:well-formed-word}
-  \end{subfigure}
-
-  \begin{subfigure}[h]{1.0\linewidth}
-    \begin{mathpar}
-  \inferrule*[Right=E-Sealed-Code]{
-    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
-    \sigma \in \sigcloss
-  }{
-    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
-  }
-  \and
-  \inferrule*[Right=E-Word]{
-    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
-  }{
-    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
-  }
-    \end{mathpar}
-    \caption{Export well-formedness}
-    \label{fig:well-formed-export}
-  \end{subfigure}
-  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
-  \label{fig:well-formed-aux}
-\end{figure}
-\end{jversion}
-\subsection{Reasonable Components}
-% Reasonability
-The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
-In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
-More precisely, we expect four things of a reasonable component:
-\begin{enumerate}[label=(\arabic*)]
-\item It checks the stack base address before performing a call.
-As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
-\item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
-\item It does not leak return and closure seals or means to retrieve them.
-This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
-There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
-\item It does not store return and closure seals or means to get them.
-By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
-\end{enumerate}
-\begin{jversion}
-  We capture these properties in 4 definitions.
-  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly as well as indirectly.
-  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then  the configuration behaves reasonable.
-  Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
-  Finally, definition~\ref{def:reasonable-component} lifts the notion of reasonability to components.
-
-  \subsubsection{Reasonable words}
-  To provide any guarantees, \stktokens{} rely on the program to no leak return seals in any way.
-  But what does it mean to ``leak'' a return seal.
-  This means that sets of seals that contain return seals as well as any means to obtain such sets cannot be leaked.
-  The following definition makes this more precise.
-\begin{definition}[Reasonable word]
-  \label{def:reasonable-word}
-  Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
-We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
-free stack $\ms_\stk$ if $n=0$
-or the following implications hold.
-  \begin{itemize}
-  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then
-    $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \mathrel{\#} \dom(\ta)$
-  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
-    \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$
-    and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all
-    $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
-  \item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$
-    and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for
-    all $\aaddr \in [\baddr,\eaddr]$
-  \end{itemize}
-\end{definition}
-According to the definition, sets of seals that contain return or closure seals are not reasonable.
-\stktokens{} need the return seals to be unique in order to work, but it does not need the closure seals.
-However, it defeats the purpose of the closure seals if they are leaked to malicious code as it allows the malicious code to fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectually allows malicious code to unseal the data capability.
-Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed in the context of.
-
-% TODO Mention sealed capabilities?
-
-A reasonable word is defined relative to a memory and a stack because for capabilities it depends on what they give access to whether they are reasonable.
-For instance, if a read capability gave access to a piece of memory that contains a set of seals with a return seal, then it would not be reasonable.
-This is why stack pointers and memory capabilities with read permission must only give access to memory with reasonable words.
-
-The definition is cyclic, but the step-index ensures that it is well-founded.
-
-\subsubsection{Reasonable pc and configuration}
-In order to define desired behaviour, we need to say specify what may happen on a machine during execution.
-The execution steps between executable configurations, so we need to define when a configuration is reasonable.
-While the step relation is defined over configurations, it is the pc that decides what instruction is executed.
-It therefore only seem natural to also define when a pc is reasonable.
-Definition~\ref{def:reasonable-pc} roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file is reasonable and the result is a reasonable configuration.
-  \begin{definition}[Reasonable pc]
-    \label{def:reasonable-pc}
-    We say that an executable capability $c =
-    ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps
-    if for any $\Phi$ such that
-    \begin{itemize}
-    \item $\Phi.\reg(\pcreg) = c$
-    \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ and
-      free stack $\Phi.\ms_\stk$ for all $r \neq \pcreg$
-    \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
-    \end{itemize}
-    We have that $\Phi$ is reasonable up to $n$ steps.
-  \end{definition}
-  With Definition~\ref{def:reasonable-word} and \ref{def:reaonable-pc} in place, we are all set to define when a configuration is reasonable.
-  \begin{definition}[Reasonable configuration]
-    \label{def:reasonable-conf}
-    We say that an execution configuration $\Phi$ is reasonable up to $n$ steps with $(\ta,\stkb,\gsigrets,\gsigcloss)$ 
-    iff for $n' \leq n$:
-    \begin{enumerate}
-    \item \label{item:guarantee-stk-base} \emph{Guarantee stack base address before call...} \\
-      If $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$
-      for some $\src{r_1}$ and $\src{r_2}$,
-      then all of the following hold:
-      \begin{itemize}
-      \item $\src{\Phi}(\src{r_\stk}) =
-        \src{\stkptr{\_,\stkb,\_,\_}}$
-      \item $r_1 \neq \rtmp{1}$
-      \item $n' = 0$ or $\Phi(\pcreg) + \calllen$ behaves reasonably up to $n'-1$ steps (Definition~\ref{def:reasonable-pc})
-      \end{itemize}
-    \item \label{item:return-seal-usage} \emph{Use return seals only for calls, use closure seals appropriately...} \\
-      If $\src{\Phi}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$,
-      then one of the following holds:
-      \begin{itemize}
-      \item $\src{\Phi}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$
-      \item $\sigma \in \gsigcloss$ and one of the following holds:
-        \begin{itemize}
-        \item $\exec{\Phi(r_1)}$ and $n' = 0$ or $\Phi(r_1)$ behaves reasonably up to $n' - 1$ steps (Definition~\ref{def:reasonable-pc}).
-        \item $\nonExec{\Phi(r_1)}$ and $n' = 0$ or $\Phi(r_1)$ is reasonable up to $n' - 1$ steps in memory $\Phi.\ms$ and free stack $\Phi.\ms_\stk$ (Definition~\ref{def:reasonable-word}).
-        \end{itemize}
-      \end{itemize}
-    \item \label{item:dont-store-private}\emph{Don't store private stuff...}\\
-      If $\src{\Phi}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$,
-      then $n' = 0$ or $\Phi.\reg(r_2)$ is reasonable in memory $\Phi.\mem$ up to $n' -1$ steps.
-    \item \label{item:dont-leak-private}\emph{Don't leak private stuff...}\\
-      If $\Phi \step[\ta,\stkb] \Phi'$, then one of the following holds:
-      \begin{itemize}
-      \item All of the following hold:
-        \begin{itemize}
-        \item $\Phi'.\reg(\pcreg) =
-          ((\perm,\lin),\baddr,\eaddr,\aaddr')$ and $\Phi.\reg(\pcreg) =
-          ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-        \item $\Phi$ does not point to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
-        \item $\Phi$ does not point to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$, $\offpc$, $\offsigma$
-        \item $n' = 0$ or $\Phi'$ is reasonable up to $n'-1$ steps
-        \end{itemize}
-        % \dominique{the following seems unnecessary: why would we allow trusted code to use a jump instruction to jump out of a trusted component?
-it can use call and xjmp for calls/returns/tailcalls}
-        % \item $\Phi'.\reg(r)$ is reasonable in memory $\Phi'.\mem$ and free stack $\Phi'.\ms_\stk$ up to $n'-1$ steps for all $r$ (including $\pcreg$)
-      \item All of the following hold:
-        \begin{itemize}
-        \item $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
-        \item $n' = 0$ or $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
-        \end{itemize}
-      \item All of the following hold:
-        \begin{itemize}
-        \item $\Phi$ points to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
-        \item $n' = 0$ or $\Phi.\reg(r)$ is reasonable in memory $\Phi.\mem$ and free stack $\Phi.\ms_\stk$ up to $n'-1$ steps for all $r \neq \pcreg$
-        \end{itemize}
-      \end{itemize}
-    \end{enumerate}
-  \end{definition}
-The four items in Definition~\ref{def:reasonable-conf} correspond to the four informal items from the introduction of this section.
-
-Item~\ref{item:dont-store-private} and Item~\ref{item:don't-leak-private} makes sure that return seals are not leaked.
-Item~\ref{item:dont-store-private} says that only reasonable words can be stored to memory.
-This means that sets of return seals or execute capabilities cannot be stored to memory by a reasonable component.
-Intuitively, a well-formed and reasonable component has its seals available in the code memory, so it can always retrieve them from the code memory.
-In other words, it is not necessary to store them in the data memory.
-Further, by making sure that seals are not stored to memory, we can allow capabilities for data memory to be passed away if there is a need for that (for instance to have a shared buffer).
-Item~\ref{item:dont-leak-private} makes sure that seals are not leaked when transferring control to another component (i.e. on security boundary crossings).
-With the component setup, there are two ways to transfer control: xjmp and call.
-In both cases, we require that all of the argument registers contain reasonable words.
-An execution configuration may need to do other operations than calling other code, and seals should no be leaked at any point during execution.
-For this reason, Item~\ref{item:don-leak-private} also says that if the next step is not a call and a jump, then the next execution configuration should also be reasonable.
-
-Item~\ref{item:guarantee-stk-base} makes sure that the stack has the correct base before a call.
-In order to not have to reason about unreasonably generated code, we also add the requirement $r_1 \neq \rtmp{1}$.
-If we allowed $r_1=\rtmp{1}$, then the call would be sure to fail as the first instruction of a call moves 42 to $\rtmp{1}$.
-Finally, this promises that the code after the call will behave reasonably.
-
-A well-formed component makes sure that a return seal is uniquely available to every call.
-This is, however, not sufficient as it does not ensure that other parts of a program don't use the return seals.
-We do not want to specify what non-call code should look like, so we just require it to not use the call seals.
-This is what Item~\ref{item:return-seal-usage} ensures.
-It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction is used to seal part of a closure which means that it should be using one of the closure seals.
-In order to construct a closure, one must seal a code capability and a data capability.
-If this is an executable capability, then it should be reasonable as a pc.
-On the other hand, if it is not an executable capability, then this must be the data part of the pair, so it should just be a reasonable word.
-
-Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle. If
-
-  \subsubsection{Reasonable component}
-  A reasonable component has the informal behavioural properties from the introduction.
-  Reasonability is captured by the previous definitions. These definitions are lifted to components by the following definition.
-  \begin{definition}[Reasonable component]
-    \label{def:reasonable-component}
-    We say that a component
-    $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$
-    is reasonable if the following hold: For all $(s \mapsto \sealed{\sigma,\vsc}) \in
-    \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$, we have that $\vsc$
-    behaves reasonably up to any number of steps $n$.
-
-    We say that a component
-    $(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ is reasonable if $\var{comp}_0$ is reasonable.
-    % \lau{Is this reasonable as a word? In that case, it should  mention the
-    %   memory and free stack.}
-    % \dominique{No: this is reasonable as a  pc.}
-  \end{definition}
-\end{jversion}
-
-In our result, we assume that adversarial components are well-formed, but not necessarily reasonable. The well-formedness assumption ensures
- that the trusted component can rely on basic security guarantees provided by the capability machine.
-For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
-The adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
-Can adversarial code call into trusted components?
-The answer to that question is yes but not with LSE and WBCF guarantees.
-Formally, adversarial code can contain the instructions that constitute a call.
-However, for untrusted code, \srccm{} will not execute those instructions as a "native call" but execute the individual instructions separately.
-The callee then executes in the same stack frame as the caller, so WBCF and LSE do not follow (for that call).
-
-We will assume trusted components, for which WBCF and LSE are guaranteed, to be both well-formed and reasonable.
-
-\subsection{Full Abstraction}
-All that is left before we state the full-abstraction theorem is to define how components are combined with contexts and executed, so that we can define contextual equivalence.
-
-\begin{figure}
-  \centering
-  \begin{equation*}
-    \inference{
-      c_{\mathrm{main},c},c_{\mathrm{main},d}  = \sealed{\sigma, c_{\mathrm{main},c}', c_{\mathrm{main},d}'} &
-      \nonExec{c_{\mathrm{main},d}'}&
-      \reg(\pcreg,\rdata) = c_{\mathrm{main},c}', c_{\mathrm{main},d}' \\
-      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
-      \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
-      \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0&
-      \range{\ms_\stk} = \{0\}&
-      \mem = \mscode \uplus \msdata \trg{\;\uplus\; \ms_\stk} \\
-      [\baddr_\stk,\eaddr_\stk] = \dom(\ms_\stk ) \mathrel{\#} (\dom(\mscode) \cup \dom(\msdata)) &
-      \overline{\var{import}} = \emptyset
-    }{
-      ((\mscode, \msdata, \overline{\var{import}}, \overline{\var{export}}, \sigrets, \sigcloss,A_\linear),  c_{\mathrm{main},c}, c_{\mathrm{main},d}) \rightsquigarrow (\mem, \reg\src{, \emptyset, \ms_\stk})
-    }
-  \end{equation*}
-  \caption{The judgement $\var{prog} \rightsquigarrow \Phi$, which defines the initial execution configuration $\Phi$ for executing a program $\var{prog}$.}
-  \label{fig:init-ec}
-\end{figure}
-
-% How a program is lifted to an executable configuration
-Given a program $\var{comp}$, the judgement $\var{comp} \rightsquigarrow \Phi$ in Figure~\ref{fig:init-ec} defines an initial execution configuration that can be executed.
-It works almost the same on \trgcm{} (conditions in \trg{red}) and \srccm{} (conditions in \src{blue}).
-On both machines a stack containing all zeroes is added, as part of the regular memory on \trgcm{} and as the free stack on \srccm{}.
-On \srccm{}, the initial stack is empty as no calls have been made.
-The component needs access to the stack, so a stack pointer is added to the register file in $\rstk$.
-On \trgcm{} this is just a linear read-write capability, but on \srccm{} it is the representation of a stack pointer.
-The entry point of the program is specified by main, so the two capabilities are unsealed (they must have the same seal) and placed in the $\pcreg$ and $\rdata$ registers.
-Other registers are set to zero.
-
-Contextual equivalence roughly says that two components behave the same no matter what context we plug them into.
-%The remaining defintions
-\begin{definition}[Plugging a component into a context]
-  \label{def:comp-context-plugging}
-  When $\var{comp'}$ is a context for component $\var{comp}$ and $\var{comp}' \bowtie \var{comp} \rightsquigarrow \Phi$, 
-  then we write $\plug{\var{comp'}}{\var{comp}}$ for the execution configuration $\Phi$.
-\end{definition}
-\begin{definition}[\trgcm{} and \srccm{} contextual equivalence]
-  \label{def:contextual-equivalence}
-  ~
-  \begin{description}
-  \item[On \srccm{}], we define that $\src{\comp_1 \sconeq \comp_2}$ iff
-  \begin{equation*}
-    \forall \src{\context} \ldotp  \wdjud[\emptyset]{\src{\context}} \Rightarrow \src{\plug{\context}{\comp_1} \sterm{\ta[,1],\stkb_1}} \Leftrightarrow \src{\plug{\context}{\comp_2} \sterm{\ta[,2],\stkb_2}}
-  \end{equation*}
-  with $\src{\ta[,i]} = \src{\dom(\comp_i.\mscode)}$.
-  \item[On \trgcm{}], we define that $\trg{\comp_1 \tconeq \comp_2}$ iff
-  \begin{equation*}
-    \forall \trg{\context} \ldotp \wdjud[\emptyset]{\trg{\context}} \Rightarrow \trg{\plug{\context}{\comp_1} \term} \Leftrightarrow \trg{\plug{\context}{\comp_2} \term}
-  \end{equation*}
-\end{description}
-where $\Phi\sterm[i]{\src{\ta,\stkb}} \text{ iff } \Phi \nstep[i]{\src{\ta,\stkb}} \halted \quad$ and $\quad \Phi\sterm{\src{\ta,\stkb}} \defeq \exists i \ldotp \sterm[i]{\src{\ta,\stkb}}$
-\end{definition}
-With the above defined, we are almost ready to state our full-abstraction, and all that remains is the compiler we claim to be fully-abstract.
-We only care about the well-formed components, and they sport none of the new syntactic constructs \srccm{} adds to \trgcm{}.
-This means that the compilation from \srccm{} components to \trgcm{} components is simply the identity function.
-% Full abstraction thm
-% \todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
-\begin{theorem}
-  \label{thm:full-abstraction}
-  For reasonable, well-formed components $\comp_1$ and $\comp_2$, we have
-  \begin{gather*}
-    \src{\comp_1} \sconeq \src{\comp_2} \quad \Leftrightarrow \quad    \src{\comp_1} \tconeq \src{\comp_2} \qedhere
-  \end{gather*}
-\end{theorem}
-Readers unfamiliar with fully-abstract compilation may wonder why Theorem~\ref{thm:full-abstraction} proves that \stktokens{} guarantees LSE and WBCF.
-Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
-This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
-If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
-In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
-In order to enforce these properties, components on \trgcm{} use \stktokens{}.
-Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
-% \todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}
-
-% \begin{itemize}
-% \item present our source language, its operational semantics (excerpts)
-% \item tell more about components, specifically well-formed according to judgement
-% \item mention our assumption of reasonability
-% \item present the full abstraction theorem.
-% \end{itemize}
-
-\section{Proving full abstraction}
-\label{sec:fa-proof}
-\input{proving-full-abstraction}
-
-\section{Discussion}
-\label{sec:discussion}
 % \begin{itemize}
@@
 
-\section{Related Work}
-
-In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.
-We do not repeat the work we discussed in Section~\ref{sec:introduction}.
-
-Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
-The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
-This is a recent, relatively mature capability machine which combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
-For simplicity, we have omitted features of CHERI that were not needed for \stktokens{} (e.g.\ local capabilities, virtual memory).
-
-Plenty of other papers enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
-This includes particularly the line of work on control-flow integrity~\citep{abadi_control-flow_2005}.
-This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static control graph information and a protected shadow stack.
-Contrary to \stktokens{}, CFI can be implemented on commodity hardware rather than capability machines.
-However, its attacker model is different, and its security goals are weaker.
-They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
-In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness saying that jumps stay within the program's static control flow graph~\cite{Abadi2005Theory}.
-Such a property ignores temporal properties and seems hard to use for reasoning.
-There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending}.
-
-More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016,juglaret_beyond_2016}.
-Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments), and we do not use per-component stacks or a trusted stack manager but a single shared stack and a decentralized calling convention based on linear capabilities.
-Both prove a secure compilation result from a high-level language which clearly implies a general form of control-flow correctness, but that result is not separated from the results about other aspects of their compiler.
-
-CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
-The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.
-They do not provide many details on this mechanism and it is, for example, not clear if and how they intend to deal with higher-order interfaces (C function pointers) or stack references shared across component boundaries.
-
-The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.
-In their study of compartmentalization primitives, \Citet{juglaret_beyond_2016} discuss the property of Secure Compartmentalizing Compilation (SCC): a variant of full abstraction that applies to unsafe source languages.
-Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
-In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
-If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
-
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-\begin{acks}
-  This research was supported in part by the ModuRes Sapere Aude
-  Advanced Grant from The Danish Council for Independent Research for
-  the Natural Sciences (FNU) and by Cost Action CA15123 EUTypes.
-  Dominique Devriese held a Postdoctoral Fellowship from the Research Foundation - Flanders (FWO) during most of this research.
-\end{acks}
-
-\bibliography{references}
-
-%% Appendix
-% \appendix
-% \section{Appendix}
-
-% Text of appendix \ldots
-
-\clearpage
-\appendix
-\input{appendix}
-\end{document}
+%%% Local Variables:
+%%% TeX-master: "paper"
+%%% End:

### 2019-06-17 popl-2019-journal-v/paper.tex
[POPLjv] Update paper preamble to not overwrite mem sat commands
[POPLjv] More work on the POPL journal version
[POPLjv] Move discussion section to new file

@@
 % Paper specific redefinitions of commands
-\renewcommand{\npair}[2][n]{#2}
 \renewcommand{\MemSeg}{\shareddom{MemSeg}}
@@
 \renewcommand{\comp}{\var{comp}}
-\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
-\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlr{F}(#3)}
 \renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
@@
 \renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
-\renewcommand{\lrr}[1][]{\lrrg[#1]{\untrusted}}
+\renewcommand{\lrr}[1][\square,\gc]{\lrrg[#1]{}}
 \renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
@@
 \renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
-\renewcommand{\lrv}[1][]{\lrvg[#1]{\untrusted}}
 \renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
@@
 
-\renewcommand{\memSat}[3][]{\memSatGeneric{#1}{#2}{#3}{}}
 \renewcommand{\stdreg}[2]{\iota^{\mathrm{std},#2}_{#1}}
@@
 \newcommand{\wdjud}[2][ ]{#1 \vdash #2}
+
+\newcommand{\erasen}[2]{\lfloor #1 \rfloor_{#2}}
 
@@
 \setlength{\belowcaptionskip}{-5pt}
+
+\renewcommand{\sectionname}{Section}
 
@@
 \label{sec:discussion}
-% \begin{itemize}
-% \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
-% \item Sharing stack references accross component boundaries is supported
-% \item Other notions of well-bracketedness (specifically one would be to allow different stacks)
-% \end{itemize}
-\subsection{Full Abstraction}
-% - Full abstraction proofs difficult
-Our formulation of WBCF and LSE using a fully abstract overlay semantics has an important advantage with respect to others.
-Imagine that you are implementing a fully abstract compiler for a high-level language, i.e.\ a secure compiler that enforces high-level abstractions when interacting with untrusted target-language components.
-Such a compiler would need to perform many things and enforce other high-level properties than just WBCF and LSE.
-
-If such a compiler uses the \stktokens{} calling convention, then the security proof should not have to reprove security of \stktokens{}.
-Ideally, it should just combine security proofs for the compiler's other functionality with our results about \stktokens{}.
-We want to point out that our formulation enables such reuse.
-Specifically, the compiler could be factored into a part that targets \srccm{}, followed by our embedding into \trgcm{}.
-If the authors of the secure compiler can prove full abstraction of the first part (relying on WBCF and LSE in \srccm{}) and they can also prove that this first part generates well-formed and reasonable components, then full abstraction of the whole compiler follows by our result and transitivity of fully abstract compilation.
-Perhaps other reusable components of secure compilers could be formulated similarly using some form of fully abstract overlay semantics, to obtain similar reusability of their security proofs.
-
-% When creating fully-abstract compilers between low-level machines, it is a big challenge to work with the exposed addresses.
-% In particular, if the compilation changes the code size of a block of code, then it may be observable and prevent full-abstraction from being proven.
-% We would argue that when a compilation reaches a phase where addresses are exposed, then the compilation should no longer change the code.
-% This does, however, pose the challenge that there might quite a few abstractions in difference between the language where addresses are hidden to a machine where they are not.
-% We propose that this challenge is solved by implementing these abstractions in a number of overlay semantics.
-% By implementing them one by one, one can deal with one abstraction at a time reducing the complexity of each necessary full-abstraction proof.
-
-% - Need to compile to a machine with enforcement mechanisms - capability machine an option
-% - Full abstraction proofs modular, so other full abstraction proofs could target \srccm{} and thus have more abstractions to work with than if \trgcm{} was the target.
-
-%\subsection{Sharing the stack}
-% ?
-\subsection{Practical Applicability}
-We believe there are good arguments for practical applicability of \stktokens{}.
-The strong security guarantees are proven in a way that is reusable as part of a bigger proof of compiler security.
-Its costs are
-\begin{itemize}
-\item a constant and limited amount of checks on every boundary crossing.
-\item possibly a small memory overhead because stack frames must be of non-zero length
-\end{itemize}
-The main caveat is that we rely on the assumption that capability machines like CHERI can be extended with linear capabilities in an efficient way.
-
-Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team as well as our own thoughts on the matter.
-As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
-Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
-Our instructions are a bit clumsy from this point-of-view because, for example, \texttt{move} or \texttt{store} will zero the source register resp.
-memory location if the value being written is linear.
-A solution for this problem could be to add separate instructions for moving, storing and loading linear registers at the cost of additional opcode space.
-Adding splice and split will also consume some opcode space.
-
-Another problem is caused by the move semantics for \texttt{load} in the presence of multiple hardware threads.
-In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
-This means that a \texttt{load} of a linear capability should behave atomically, similar to a primitive compare-and-swap instruction.
-This is in principle not a problem except that atomic instructions are significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
-When using \stktokens{}, loads of linear capabilities happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
-Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, well-engineered caching could perhaps reduce the high overhead of atomic loads of linear capabilities.
-% If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
-The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
+\input{discussion}
 
 \section{Related Work}
-
 In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.

### 2019-06-17 popl-2019-journal-v/discussion.tex
[POPLjv] Add TODO comments to discussion and related work

@@
 % ?
+
+\begin{jversion}
+  A compiler is fully abstract when it enforces all high-level properties of a high-level language which means that the high-level language that dictates what we need to enforce.
+  In other words to figure out what properties we want, we need to figure out what high-level language we want.
+  % TODO is this all we have to say about high-level properties? Maybe something about address hiding and the potential issues. Also other notions of control-flow, e.g. call/CC
+  % TODO say something about horizontal vs vertical composition.
+
+  % TODO new paragraph: conditional full-abstraction
+
+  % TODO new paragraph: Machine checked proof. MANY details, any future work should be machine checked.
+
+  % TODO new subsection on construction of overlay semantics:
+  % If we don't, we will have to come up with new syntax. Further, say we had
+  % introduced new instruction, how would we do the back translation? We would not
+  % be able to tell whether a series of instructions came from a call or a series of
+  % instructions that just look like a call. With this approach everything that
+  % looks like a call is a call. The back translation is simply the identity making
+  % things a lot easier.
+
+\end{jversion}
+
 \subsection{Practical Applicability}
@@
 
+% TODO passing pointers for stack allocated memory is not possible with linear stack
+
 %%% Local Variables:

### 2019-06-20 popl-2019-journal-v/discussion.tex
[POPLjv] Resolve some discussion todos

--- 2019-06-17 popl-2019-journal-v/discussion.tex
@@
 % ?
+
+\begin{jversion}
+  A compiler is secure when it enforces the properties of high-level languages which begs the question what properties should we enforce.
+  When it comes to fully-abstract compilation, then the answer is that all the properties of the high-level language should be enforced which means that the real question is what high-level language we would like.
+  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/CC, then we need to come up with a different enforcement scheme.
+  Further, many high-level languages have exceptions which is yet another form of control-flow which is also not supported by \stktokens{}.
+  This goes to show how we must consider what high-level language we want in order to answer the question of what properties we must enforce to get a fully-abstract compiler.
+  % Lau: Say something about address hiding?
+
+  % Say something about horizontal vs vertical composition.
+  Fully-abstract compilers compose vertically which means that if we defined a compiler phase that targets \srccm{}, then we could use the full-abstraction theorem immediately.
+  However, fully-abstract compilers do not extend vertically in the sense that if we add a new feature in the target and source language, then the full-abstraction theorem does not follow immediately.
+  Intuitively, this is the case because the new feature may have an unforeseen interaction with existing features that breaks a desired property.
+  In order to enforce more language properties on capability machines, it is reasonable to assume that new capability primitives may be necessary.
+  \stktokens{}, for instance, requires linear capabilities, and \citet{skorstengaard_reasoning_2017} uses local capabilities to enforce WBCF and LSE.
+  In order to implement address-hiding (i.e.\ have references that do not expose the address they actually point to), it is reasonable to believe that one would need a new capability primitive.
+  After all, the capabilities we use here do not hide the addresses in any meaningful way, so a new ``address hiding'' capability would be necessary to enforce this property.
+  If we extended \trgcm{} and \srccm{} with new such capabilities, then we would have to redo the entire full-abstraction proof.
+
+  % new paragraph: conditional full-abstraction
+  Our full-abstraction theorem, Theorem~\ref{thm:full-abstraction}, is not pure full-abstraction as it requires the components to be reasonable and well-formed.
+  In other words, if we were to define a compiler phase that targets \srccm{}, then we would also have to show that every program it generates is well-formed and reasonable in order to use the full-abstraction result.
+  However, even though the well-formed and reasonable conditions are fairly simple, it is too late to ensure them at this point of the compilation.
+  For instance, reasonability require that return seals are not passed to the adversary in calls.
+  In order to ensure this with \stktokens{}, we would have to add checks to the calling convention that ensures that seals are not passed as arguments in calls.
+  In other words, we would need to add protection to the trusted program against itself which of course shouldn't be necessary.
+  The calling convention should only include protection schemes that protects against untrusted programs.
+  The compiler should generate programs that are actually trustworthy, so we do not have to add protection against the generated programs.
+  However, when we consider \srccm{} components, we have already committed to a program consisting of low-level instructions.
+  The correct place to ensure that our programs are well-formed and reasonable is when we actually generate the instructions and commit to the low-level code which is the phase that goes from the high-level language (or some intermediate representation) to a low-level machine like \srccm{}.
+
+  % TODO new paragraph: Machine checked proof. MANY details, any future work should be machine checked?
+
+  % TODO new subsection on construction of overlay semantics:
+  % If we don't, we will have to come up with new syntax. Further, say we had
+  % introduced new instruction, how would we do the back translation? We would not
+  % be able to tell whether a series of instructions came from a call or a series of
+  % instructions that just look like a call. With this approach everything that
+  % looks like a call is a call. The back translation is simply the identity making
+  % things a lot easier.
+
+\end{jversion}
+
 \subsection{Practical Applicability}
@@
 
+\begin{jversion}
+Programs that use a C-like calling convention often allow programs to pass a stack reference in a call either to provide a reference for a piece of memory to work on or to indicate where the result of the call should be placed.
+\stktokens{} does not support this as it expects to get a stack pointer for the entire stack it handed out.
+However, if we have allocated part of the stack for a different purpose and handed out a reference for it, then the stack pointer cannot reference the same memory as the stack pointer is a linear capability.
+However, while it is not possible to allocate a reference for the stack, one can easily imagine that the stack could still be used to pass and return values.
+\end{jversion}
+
 %%% Local Variables:

### 2019-06-20 popl-2019-journal-v/discussion.tex
[POPLjv] Discussion of overlay semantics and back-translation

@@
   % TODO new paragraph: Machine checked proof. MANY details, any future work should be machine checked?
-
-  % TODO new subsection on construction of overlay semantics:
-  % If we don't, we will have to come up with new syntax. Further, say we had
-  % introduced new instruction, how would we do the back translation? We would not
-  % be able to tell whether a series of instructions came from a call or a series of
-  % instructions that just look like a call. With this approach everything that
-  % looks like a call is a call. The back translation is simply the identity making
-  % things a lot easier.
-
+  One challenge in full-abstraction proofs is to relate the translation of program to the program it was translated from.
+  Such a relation is often expressed as a back-translation  \citep{devriese_modular_2017}, i.e.\ a translation from the target language to the source language.
+  When we use an overlay semantics, the back-translation becomes trivial because the source and target language are syntactically the same, so the identity can be used as the back-translation.
+  If we have native call and return instructions in the source language, then the source language would be different from the target language, and we would have to use a non-trivial back-translation.
+  Specifically, the back-translation would need to distinguish sequences of instructions that is the translation of a call from sequences of instructions that just look like a call.
+  With overlay semantics, this is not a concern because everything that looks like a call is interpreted as a call.
 \end{jversion}

### 2019-07-02 local-capability-machine/local-capability-machine.tex
Corrected mistakes with configuration with halted instead of memory. Added some implicit text for lemma added to toplas paper.
typo in local cap tech report
typo
local cap tech report naming
Fix mclear macro
Thanks, Aina.

@@
         move r_t2 pc
-        lea r_t2 ... // ... is the offset to end
+        lea r_t2 ... // ... is the offset to continue
         move r_t3 pc
@@
         jnz r_t2 r_t1
+        lea pc ... // ... if the offset to end
+continue:
         store r_t 0
         lea r_t 1
-        plus r_t1 r_t1 1
+        minus r_t1 r_t1 1
         jmp r_t3

### 2019-07-04 toplas/paper.tex
[TOPLAS] Disable toplas command and environment
Update my affiliation.
[TOPLAS] Update malloc specification

@@
 \definecolor{Blue}{RGB}{92,26,143}
-\newenvironment{toplas}%{}{}
-    {\color{OliveGreen}}{}
+\newenvironment{toplas}{}{}
+%    {\color{OliveGreen}}{}
 \newenvironment{toplassug}
@@
 
-\newcommand{\itoplas}[1]%{#1}
-    {{\color{OliveGreen} #1}}
+\newcommand{\itoplas}[1]{#1}
+%    {{\color{OliveGreen} #1}}
 \newcommand{\itoplassug}[1]{}
@@
 %  \position{Position1}
-  \department{iMinds-DistriNet}              %% \department is recommended
-  \institution{KU Leuven}            %% \institution is required
-  \streetaddress{Celestijnenlaan 200 A}
-  \city{Leuven}
-  \postcode{3001}
+  \department{Vakgroep Computerwetenschappen}              %% \department is recommended
+  \institution{Vrije Universiteit Brussel}            %% \institution is required
+  \streetaddress{Pleinlaan 2}
+  \city{Elsene}
+  \postcode{1050}
   \country{Belgium}
 }
-\email{dominique.devriese@cs.kuleuven.be}          %% \email is recommended
+\email{dominique.devriese@vub.be}          %% \email is recommended
 
@@
 % \institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
-% \and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}
+% \and Vrije Universiteit Brussel \\ \email{dominique.devriese@vub.be}}
 
@@
           \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land{} \\
-          \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land{} \\
+          \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\arraycolsep=0pt\array[t]{l}\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\\\update{r_1}{((\rwx,\glob),b',e',b')}\\\update{r_\var{t1},r_\var{t2},r_\var{t3}}{0,0,0} \land{}\endarray  \\
           \qquad \quad \size - 1 = e'-b'
@@
 
-For convenience, we also require that malloc returns the non-argument registers
-of the register-file unchanged after the call. This allows the caller to keep
-private capabilities in the register file, without having to protect them by
-storing them in a private stack frame.
+For convenience, we also require that malloc returns the non-argument registers (except registers for malloc internal computations) of the register-file unchanged after the call.
+This allows the caller to keep private capabilities in the register file, without having to protect them by storing them in a private stack frame.
 

### 2019-07-12 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add sketch of the memory to world explanation

@@
 \srccm{}s memory is split in three: heap, free stack and encapsulated local stack frames.
+The three kinds of memory and their relative position to each other are sketched in Figure~\ref{fig:mem-sketch}.
+Note how the encapsulated local stack frames are sparsely laid out in the stack with free stack in between.
+The free stack in between stack frames may be used by untrusted code for their local stack frames.
+Untrusted stack frames are not protected by \srccm{}, but this does not prevent untrusted code from securing their own stack frames.
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+
+  \scope
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+  \endscope
+  \draw (-1,0) node {};
+  \draw (-1,12) node {};
+
+  \draw[pattern=dots,draw=none] (0,0) rectangle (4.5,2);
+  \draw[pattern=dots,draw=none] (0,9.5) rectangle (4.5,11);
+
+  \draw[pattern=north west lines] (0,2) rectangle (4.5,4);
+  \draw[pattern=crosshatch] (0,4) rectangle (4.5,5);
+  \draw[pattern=north west lines] (0,5) rectangle (4.5,6);
+  \draw[pattern=crosshatch] (0,6) rectangle (4.5,7.5);
+  \draw[pattern=north west lines] (0,7.5) rectangle (4.5,8.5);
+  \draw[pattern=crosshatch] (0,8.5) rectangle (4.5,9.5);
+
+  \draw (5.75,2.25) node[anchor=west] {\tiny $\stkb$};
+  \draw (5.75,2.25) edge[-latex] (4.75,2.25);
+  \draw (5,9.5) edge[-latex] (5,8);
+  \draw [decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt},yshift=0pt] (-0.25,9.5) -- (-0.25,2) node [black,midway,xshift=-.75cm,yshift=0.05cm,rotate=45] {\footnotesize stack};
+
+  \matrix [draw,below left] at (current bounding box.north west) {
+    \node[pattern=dots,draw,label=right:{\tiny Heap ($\ms$)}] {}; \\
+    \node[pattern=north west lines,draw,label=right:{\tiny Free stack ($\ms_\stk$)}] {}; \\
+    \node[pattern=crosshatch,draw,label=right:{\tiny Encapsulated stack frame ($\stk$)}] {}; \\
+};
+
+  \end{tikzpicture}
+  \caption{A sketch of how heap, encapsulated stack and free stack are laid out in memory.
+    The encapsulated stack frames and the free stack constitutes the stack.
+    The encapsulated stack frames may have free stack in between them where stack frames of non-trusted code may reside.
+    The stack grows downwards in memory with $\stkb$ as the top address (and the base address of the free stack capability).}
+  \label{fig:mem-sketch}
+\end{figure}
+
 In order to model the three kinds of memory, we simply have three sub-worlds.

### 2019-07-12 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Move legend of memory sketch
[POPLjv] Rewrite spatial introduction

@@
 
-  \matrix [draw,below left] at (current bounding box.north west) {
+  \matrix [draw,below right] at (current bounding box.north east) {
     \node[pattern=dots,draw,label=right:{\tiny Heap ($\ms$)}] {}; \\
@@
 It is only the spatial owned regions that can be referenced by a capability.
+Even though only spatial owned regions can be referenced by linear capabilities, we still have unclaimed spatial regions that specify what part of memory looks like.
+When there is an unclaimed spatial region, then there may be a compatible world  that claims the ownership of that spatial region which means that the memory can be referenced (we return to the notion of compatible worlds w.r.t.
+ownership later).
 Specifically, we add tags $\spatial$ and $\spatialo$ to the spatial regions:
@@
 
-Given a world, we sometimes want to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
+Given a world, we want to be able to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
 This is expressed by stripping the world of all its ownership, so the only active regions are the $\pure$ regions.

### 2019-07-12 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Make purPart function def more concise

@@
     \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
-    \purePart{W_\var{heap}} &\defeq
+    \purePart{W_\var{heap}} &\defeq \lambda r\ldotp
                        \begin{cases}
-                         W_{\var{heap}}(r) & \text{if } W_\var{heap}(r) = (\pure,\var{sm})\\
-                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatial,\var{sm})\\
                          (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
-                         \revoked & \text{if } W_\var{heap}(r) = \revoked\\
+                         W_{\var{heap}}(r) & \text{otherwise}
                        \end{cases}\\
-    \purePart{W_\var{priv}} &\defeq
+    \purePart{W_\var{priv}} &\defeq \lambda r\ldotp
                        \begin{cases}
-                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatial,\var{sm}),\opc)\\
                          ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\opc)\\
-                         (\revoked,\opc) & \text{if } W_\var{priv}(r) = (\revoked,\opc)\\
+                         W_\var{priv}(r) & \text{otherwise}
                        \end{cases}\\
-    \purePart{W_\var{free}} &\defeq
+    \purePart{W_\var{free}} &\defeq \lambda r\ldotp
                        \begin{cases}
-                         W_{\var{free}}(r) & \text{if } W_{\var{free}}(r) = (\pure,\var{sm})\\
-                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatial,\var{sm})\\
                          (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
-                         \revoked & \text{if } W_{\var{free}}(r) = \revoked\\
+                         W_{\var{free}}(r) & \text{otherwise}
                        \end{cases}

### 2019-07-12 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add a word on linear capabilities being affine

@@
 % StkTokens designate seals for certain purposes, return, closure, etc
-In order to guarantee well-bracketed control flow and local-state encapsulation, \stktokens{} designates seals for return capabilities and closure seals and relies on designated seals only being used for their intended purpose.
+In order to guarantee well-bracketed control flow and local-state encapsulation, \stktokens{} designates seals for return capabilities and closures and relies on designated seals only being used for their intended purposes.
 The designation of return seal is a system invariant that is modelled in the world as a \textit{seal interpretation function}.
@@
 The seal interpretation function should be an invariant, so we need to add it to a type of region that is invariant.
-As we will see, shared regions are invariant, so we add the seal interpretation function as follows
+As we will see in Section~\ref{subsubsec:ft-and-revocation}, shared regions are invariant, so we add the seal interpretation function as follows
 \begin{multline*}
@@
 
-We will also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it i unambiguous.
+We will also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it is unambiguous.
 
@@
 % Revocation of spatial regions. 'revocation' of linear capabilities.
-All regions are future regions of themselves, so regions are allowed to stay unchanged.
-The $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
-% TODO describe why "Spatial -> spatialo" ~ affine or at least why we have it?
+The first rule says that all regions are future regions of themselves, so regions are allowed to stay unchanged.
+The second rule says that $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
+The third and final rule says that $\spatial$ regions can become $\spatialo$ which indicates that the linear capabilities are affine.
+Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes them affine.
+This is modelled in the world by allowing future worlds to pickup ownership.
+% describe why "Spatial -> spatialo" ~ affine or at least why we have it?
 \todo[inline]{Lau: explain why $\spatialo \future \spatial$ gives us affine capabilitites.
-This is claimed in the technical report, but I don't see how this is different from revoking $\spatial$ regions and adding a new region with the future world relation.}
+This is claimed in the technical report, but I don't see how this is different from revoking $\spatial$ regions and adding a new region with the future world relation...
+I have written something, but it does not feel like the story really holds.}
 \begin{figure}[htb]
@@
   \begin{mathpar}
+  \inferrule{ r \in \Regions \cup \Regionh }{ r \future r}
+  \and
   \inferrule{ }{ \revoked \future (\spatial,\_)}
@@
   \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
-  \and
-  \inferrule{ r \in \Regions \cup \Regionh }{ r \future r}
 \end{mathpar}

### 2019-07-12 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Address comments

@@
 
-With the future region relation in place, we define the future world relation as follows:
-\begin{mathpar}
-  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r) }
-            { W' \future W }
-\end{mathpar}
+With the future region relation in place, we define the future world relation as follows: For worlds $W$ and $W'$,
+\[
+  W' \future W \text{ iff } \left\{
+    \array{l}
+    \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\
+    \quad\exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
+    \qquad \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r)
+     \endarray
+  \right.
+\]
 The relation says that each of the three worlds must be an extension of the past world and each of the existing region must have a future region.
@@
 Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
-which provide sufficient structure for us to solve recursive domain equations.
 In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.
@@
 \end{definition}
-The $\uplus$ world join does not guarantee that the result is a sensible world with respect to authority.
-However, it does not guarantee that a join is a sensible memory specification.
+The $\uplus$ world join does not guarantee that the result is a sensible world with respect to authority or memory specification.
 In \sectionname~\ref{subsubsec:mem-sat}, we define memory satisfaction which also acts as a well-formedness judgement.
@@
 When we view the world as a specification of authority, then the world join need to respect the region ownership.
-That is, when we join the authority of two worlds, then the ownership of the two worlds should be completely separate.
+That is, when we join the authority of two worlds, then the ownership of the two worlds should not overlap.
 This is expressed by the $\oplus$ operator.
@@
 Note also that this picture is further complicated by our usage of non-authority-carrying $\spatial$ regions.
-They are really only there in a world $W$ as a shadow copy of a $\spatialo$ region in another world $W'$ that $W$ will be combined with.
+They are really only in a world $W$ as a shadow copy of a $\spatialo$ region in another world $W'$ that $W$ will be combined with.
 The shadow copy is used for specifying when a memory satisfies a world: the memory should contain all memory ranges that anyone has authority over, not just the ones whose authority belongs to the memory itself.
@@
 Memory satisfaction is split into four definitions.
-At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory triple, $\ms_S$, $\ms_\stk$, and $\stk$, from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
+At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory triple, $\ms_S$ (heap), $\ms_\stk$ (free stack), and $\stk$ (stack frames), from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
 The top-level definition of memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
@@
       \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
-      \qquad W = \oplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
+      \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
       \qquad \forall r \in \dom(\activeReg{W_\var{stack}}),n' < n \ldotp \\

### 2019-07-12 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Typos

@@
 \end{definition}
-The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible, but when we use a world, we will always make sure that it is sensible by using memory satisfaction.
+The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible (e.g., the regions are not overlapping), but when we use a world, we will always make sure that it is sensible by using memory satisfaction.
 

### 2019-07-15 popl-2019-journal-v/paper.tex
[POPLjv] Add TODO comments to discussion and related work
[POPLjv] Proofreading

@@
+% TODO do we say something about linear pcs and why they don't make sense? (related to def:reasonable-pc)
+
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
@@
   The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all project information about capabilities.
-The $\tgetb{}{}$ instruction projects the base address of the range of authority, and $\tgete{}{}$ projects the end address of the range of authority.
+  The $\tgetb{}{}$ and $\tgete{}{}$ instructions, respectively, project the base and end address of the range of authority.
   The linearity of a permission is projected with $\tgetlin{}{}$, and finally the permission is projected with $\tgetp{}{}$.
-The instructions $\tgetb{}{}$ and $\tgete{}{}$ also works on sets of seals.
-The instruction $\tisptr{}{}$ gives the type of a word which allows programs to be defensive in the sense that they can check whether a word is a capability before they use the word as one.
+  The instructions $\tgetb{}{}$ and $\tgete{}{}$ also work on sets of seals.
+  The instruction $\tisptr{}{}$ returns an integer representation of the type of a word which allows programs to be defensive in the sense that they can check whether a word is a capability before they use it.
   \trgcm{} also has arithmetic instructions $\tplus{}{}{}$, $\tminus{}{}{}$, and $\tlt{}{}{}$.
@@
   The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set of seals) to the base address of the range of authority (or range of seals).
-This instruction makes it possible to work relatively to the base address of a capability rather than with fixed addresses.
-While this instruction is not strictly necessary as it can be emulated with other instruction, it still seems useful in the case where you receive capabilities from an untrusted source.
-Finally, we have the $\trestrict{}{}$ instruction which  restricts the permission of a capability but only according to the $\le$ relation.
+  This instruction makes it easy to work relatively to the base address of a capability.
+  This instruction is not strictly necessary as it can be emulated with other instructions.
+  Finally, we have the $\trestrict{}{}$ instruction which restricts the permission of a capability according to the $\le$ relation.
 \end{jversion}
@@
   To motivate the necessity of an encapsulation mechanism like sealing, consider the following example.
-  We want to transfer control to code that we do not trust, and we want to give them the means to return to us.
+  We, a trusted piece of code, want to transfer control to code that we distrust, and we want to give them the means to return to us.
   That is, we need to give them a return capability.
   If we did not have an encapsulation mechanism, our only option would be to give them an executable capability for the address we want them to return to.
-  The code we give the return capability could use it as intended, but it could also manipulate the return capability to point to a different point and return to an unintended point of our code.
-  Further in order to be able to retrieve our capabilities from before transferring control, we would have to store them somewhere accessible from the return capability.
-  However, the untrusted code would have access to all this through the return capability because it does not give less authority to the untrusted code than it does to us.
-  For this reason, any reasonable capability machine must have an encapsulation mechanism to enable programs to make boundaries between security domains.
-
-  In the example, if we use sealing, we can seal the return capability which does two things.
-  It prevents the untrusted code from changing the target of the return capability forcing them to return to the point we intended.
-  It also prevents the untrusted code from reading our capabilities.
-  All in all, this means that we can transfer control to another domain without giving up our capabilities or handing them over.
-
-  \trgcm{} does not have a direct unsealing instruction, but it is still possible to do some amount of unsealing.
-  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains the $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
+  The untrusted code could use the return capability as intended, but it could also manipulate and make it point to a different address of our code.
+  Jumping to such a capability would cause an the program to execute in a way we did not intend for it.
+  Further in order to be able to retrieve our capabilities from before transferring control, we would have to store the capabilities somewhere accessible from the return capability.
+  However, the untrusted code would have access to all this through the return capability because it gives the same authority to us as the untrusted code.
+  This is why, any reasonable capability machine must have an encapsulation mechanism to allow programs to make boundaries between security domains.
+
+  In the example, we could seal the return capability to establish a boundary.
+  Specifically, it would prevent the untrusted code from changing the target of the return capability forcing them to return to the point we specified.
+  It would also prevent the untrusted code from reading our capabilities.
+  All in all, this means that we can transfer control to untrusted code without giving up our capabilities or handing them over.
+
+  \trgcm{} does not have a direct unsealing instruction, but it is still possible to emulate a limited unsealing mechanism.
+  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
   Now assume part of the code would like to have access to $w$ in $\rdata$.
@@
   However, given a sealed executable capability $\sealed{\sigma,c}$ and a set of seals that contains $\sigma$, we can still construct the data part of the sealed capability pair.
-  This means that we can execute the code the capability points to in the presence of any data we can construct.
-
-  Sealing is meant for encapsulation, but it gives us very weak encapsulation properties if seals are not private as the unsealing example illustrates.
-  For this reason, it is important that the system is setup, so each component has access to seals that are unique to that component.
+  This means that we can execute the code the capability points to together with data that it was not intended to be used with.
+
+  Sealing is meant for encapsulation, but it relies on seals being kept private as illustrated by the unsealing emulation.
+  For this reason, it is important that the system is initialised such that each component has access to unique seals.
   We return to this in \sectionname~\ref{sec:well-form-reas}.
@@
 \subsection{Decoding and encoding functions}
-The operational semantics of the capability machine uses $\decInstr{}$ to decode instructions.
-We also assume an $\encInstr{}$ which is used to put programs onto the capability machine.
+The operational semantics of the capability machine uses the function $\decInstr{}$ to decode instructions.
+We also assume a function $\encInstr{}$ to make it easy to specify programs in terms of instructions.
 Rather than specifying a decode function and an encode function, we assume that they are given with certain properties.
@@
 \[
-\forall c \in \Cap \ldotp \decInstr{c} = \fail
+\forall c \in \Caps \ldotp \decInstr{c} = \fail
 \]
@@
 The machine also assumes decode and encode functions for permissions $\decPerm{} : \Perm \fun \ints$ and $\encPerm{} : \ints \fun \Perm$.
-We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$ and surjective. For $\encPerm{}$, we assume it is surjective and that it does not encode anything to the error value -1, i.e.
+We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$ and surjective.
+For $\encPerm{}$, we assume it is surjective and that it does not encode anything to the $\tgetp{}{}$ error value -1, i.e.
 \[
@@
 
-Finally, the machine uses an encode function for word types $\encType{}$ in the interpretation of $\tisptr{}{}$.
-This function encodes each kind type of word as an integer.
+Finally in the interpretation of $\tisptr{}{}$, the machine uses an encode function for word types $\encType{}$.
+This function encodes each kind of word as an integer.
 This is very much like the previous functions.
@@
 The overlay semantics also has a type encoding function which must encode the new words, so they correspond to their representation on this machine.
-For instance, the encoding of the type of a stack-pointer in the overlay semantics should be the same as the encoding of a capability on this machine.
-All the new words correspond to a word on this machine, and their encodings should match.
+For instance, the encoding of stack-pointer in the overlay semantics should be the same as the encoding of a capability on this machine.
+All the new words in the overlay semantics correspond to a word on this machine, and their encodings should match.
 \end{jversion}
@@
 % - Address must be in data memory
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
-The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
-
 \begin{jversion}
@@
 \begin{jversion}
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on system guarantees.
+The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
 The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
@@
 The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
-The \textsc{Main} rule requires the component to have a main pair picked from the exports.
+The \textsc{Main} rule requires the component to have an entry point in terms of a main pair from the exports.
 Further, the component should be a base component that satisfies judgement.
-The \textsc{Base} judgement makes sure that the component has a bit of structure.
+The \textsc{Base} judgement ensures sure that the component has a bit of structure.
 Roughly speaking, the \textsc{Base} rule requires the following:
@@
 \item \textit{Code and data memory are disjoint.}
-\item \textit{Code memory is enclosed in a memory segment $\ms_\var{pad}$ that contains zero's.} This prevents the memory of two code memories from being spliced together.
-\item \textit{Data memory can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.} The component judgement ensure that the linear address space of a component is at most referenced by one capability.
+\item \textit{Code memory is padded with zeros ($\mspad$).} This prevents the memory of two code memories from being spliced together.
+\item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.} The component judgement ensure that the linear address space of a component is at most referenced by one capability.
 This means that any capability for the linear address space is unique.
@@
   \begin{itemize}
-  \item \textit{The code memory is disjoint from the trusted address space $\ta$ and in that case there are no return seals.} In this case, the component contains untrusted code.
-We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals.
-  \item \textit{The code memory is part of the trusted address space $\ta$.} In this case, the component contains trusted code.
+  \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code.
+We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to seals that could be used to protect calls.}.
+  \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code.
 We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
   \end{itemize}
-\item \emph{The data memory is disjoint from the trusted address space $\ta$.} We are not looking to execute data, so the trusted addresses never include the data memory addresses.
+\item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
 \item \textit{The code memory satisfies the component-code judgement (\figurename~\ref{fig:well-formed-code}).}
-\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).
-In particular, this must be with respect to the linear addresses assigned to this address.
-}
-\item \textit{All the exports satisfies the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
+\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address.
+\item \textit{All the exports satisfy the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
 \end{itemize}
 
-\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ the words that are well-formed in a component's data memory.
+\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ which specifies the words that are well-formed in a component's data memory.
 The judgement is defined by three rules: \textsc{W-Data}, \textsc{W-Capability}, and \textsc{W-Sealed-Capability}.
@@
 In particular, this means that sealed capabilities for the data memory are acceptable exports which allows components to export the data part of a sealed capability pair.
-The \textsc{E-Sealed-Code} rule allows the code part of a sealed capability pair to be export.
+The \textsc{E-Sealed-Code} rule allows the code part of a sealed capability pair to be exported.
 In particular, it allows a capability with read-execute permission and its range of authority within the component's code address space to be exported when it has been sealed with a closure seal.
-Together, these two rules allow components to have closures that they export.
+Together, these two rules allow components to export closures.
 
@@
 One judgement considers the entire memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ while the other considers the contents of each code memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
-Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each of its words in isolation.
+Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation.
 The well-formedness of code memory depends on whether seals for calls are located in the place we expect them to be located.
@@
 \end{definition}
-This definition says that if code memory contains an instruction that may be part of a call, then either it is a part of a call or there is some other instruction that witnesses the fact that it is not part of a call.
+Definition~\ref{def:no-hidden-calls} says that if code memory contains an instruction that may be part of a call, then either it is a part of a call or there is some other instruction that witnesses the fact that it is not part of a call.
 Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set of seals available.
@@
   We capture these properties in 4 definitions.
-  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly as well as indirectly.
-  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then  the configuration behaves reasonable.
+  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly or indirectly.
+  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then  the configuration behaves reasonably.
   Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
@@
   To provide any guarantees, \stktokens{} rely on the program to no leak return seals in any way.
-  But what does it mean to ``leak'' a return seal.
-  This means that sets of seals that contain return seals as well as any means to obtain such sets cannot be leaked.
+  But what does it mean to ``leak'' a return seal?
+  It means that sets of seals that contain return seals as well as any means to obtain such sets cannot be leaked.
   The following definition makes this more precise.
@@
   \label{def:reasonable-word}
-  Take a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
+  Take a set of trusted addresses $\ta$ and sets of return and closure seals $\gsigrets$ and $\sigcloss$.
 We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
@@
     $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
-  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \mathrel{\#} \dom(\ta)$
+  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \mathrel{\#} \ta$
   \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
@@
 According to the definition, sets of seals that contain return or closure seals are not reasonable.
-\stktokens{} need the return seals to be unique in order to work, but it does not need the closure seals.
+\stktokens{} rely on return seals to be unique in order to work, but it does not rely on closure seals.
 However, it defeats the purpose of the closure seals if they are leaked to malicious code as it allows the malicious code to fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectually allows malicious code to unseal the data capability.
-Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed in the context of.
+Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed with.
 
@@
 
-A reasonable word is defined relative to a memory and a stack because for capabilities it depends on what they give access to whether they are reasonable.
-For instance, if a read capability gave access to a piece of memory that contains a set of seals with a return seal, then it would not be reasonable.
+Whether a capability is reasonable depends on what it gives access to.
+This is why the word reasonability is defined relative to the memory and stack.
+For instance, if a read capability gives access to a piece of memory that contains a set of seals with a return seal, then it would not be reasonable.
 This is why stack pointers and memory capabilities with read permission must only give access to memory with reasonable words.
-
 The definition is cyclic, but the step-index ensures that it is well-founded.
@@
 \subsubsection{Reasonable pc and configuration}
-In order to define desired behaviour, we need to say specify what may happen on a machine during execution.
-The execution steps between executable configurations, so we need to define when a configuration is reasonable.
+In order to define desired behaviour, we need to specify what may happen on a machine during execution.
+An execution steps between executable configurations, so we need to define when a configuration is reasonable.
 While the step relation is defined over configurations, it is the pc that decides what instruction is executed.
-It therefore only seem natural to also define when a pc is reasonable.
-Definition~\ref{def:reasonable-pc} roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file is reasonable and the result is a reasonable configuration.
+Therefore, it only seems natural to also define when a pc is reasonable.
+Definition~\ref{def:reasonable-pc} roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file are reasonable and the result is a reasonable configuration.
   \begin{definition}[Reasonable pc]
@@
   \end{definition}
-  With Definition~\ref{def:reasonable-word} and \ref{def:reaonable-pc} in place, we are all set to define when a configuration is reasonable.
+  With Definition~\ref{def:reasonable-word} and \ref{def:reasonable-pc} in place, we are all set to define when a configuration is reasonable.
   \begin{definition}[Reasonable configuration]
@@
 
-Item~\ref{item:dont-store-private} and Item~\ref{item:don't-leak-private} makes sure that return seals are not leaked.
+Item~\ref{item:dont-store-private} and Item~\ref{item:dont-leak-private} make sure that return seals are not leaked.
 Item~\ref{item:dont-store-private} says that only reasonable words can be stored to memory.
@@
 An execution configuration may need to do other operations than calling other code, and seals should no be leaked at any point during execution.
-For this reason, Item~\ref{item:don-leak-private} also says that if the next step is not a call and a jump, then the next execution configuration should also be reasonable.
+For this reason, Item~\ref{item:dont-leak-private} also says that if the next step is not a call and a jump, then the next execution configuration should also be reasonable.
 
 Item~\ref{item:guarantee-stk-base} makes sure that the stack has the correct base before a call.
-In order to not have to reason about unreasonably generated code, we also add the requirement $r_1 \neq \rtmp{1}$.
+In order to not have to reason about unreasonably generated code, we also add the requirement $r_1 \neq \rtmp{1}$ before calls.
 If we allowed $r_1=\rtmp{1}$, then the call would be sure to fail as the first instruction of a call moves 42 to $\rtmp{1}$.
@@
 This is what Item~\ref{item:return-seal-usage} ensures.
-It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction is used to seal part of a closure which means that it should be using one of the closure seals.
+It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction seals part of a closure and uses a closure seal.
 In order to construct a closure, one must seal a code capability and a data capability.
@@
 On the other hand, if it is not an executable capability, then this must be the data part of the pair, so it should just be a reasonable word.
-
 Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle. If
@@
     We say that a component
-    $(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)$
+    \[(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\]
     is reasonable if the following hold: For all $(s \mapsto \sealed{\sigma,\vsc}) \in
@@
 
+% TODO Relate full-abstraction to other notions of secure compilation
+
 % \begin{acks}                            %% acks environment is optional

### 2019-07-15 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Proof read world presentation

@@
 The difficult part is to define when components are related.
-In the next section, we give an informal overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
+In the next section, we give an overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
+\begin{jversion}
+In this presentation, we try to include most of the details.
+We refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
+\end{jversion}
 
@@
 To accommodate such conditional relatedness, we construct the relation as a step-indexed Kripke logical relation with recursive worlds.
-\begin{jversion}
-In this presentation, we try to include most of the details.
-We refer to the conference version of this paper for a presentation\lau{insert reference} with fewer details and emphasis on the intuition.
-\end{jversion}
 
@@
 The possible contents of the memory also influences what the world looks like.
-For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} are modelled by the worlds.
-In order to related \trgcm{} and \srccm{}, we model all the features of \srccm{} in the worlds which means we have to model:
+For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds.
+In order to relate \trgcm{} and \srccm{}, we model all the features of \srccm{} in the world which means we have to model:
 \begin{itemize}
@@
 Linearity is modelled by adding ownership to certain parts of the world, so capabilities can take ownership of part of the memory thus ensuring that they are the sole reference.
-The call stack is modelled by the memory satisfaction (the relation that decides whether two memories are related in the world) by ensuring that the memory is actually shaped like a stack
-Finally, the seals are modelled by seal invariants that make sure that seals only are used on permitted sealables.
-In the following, we present the world and go into details about how each of the four points are modelled.
+The call stack is modelled by the memory satisfaction (the relation that decides whether two memories are related in the world) by ensuring that the memory is actually shaped like a stack.
+Finally, the seals are modelled by seal invariants that make sure that seals are only used on permitted sealables.
+In the following, we present the world and go into details about how each of the four features are modelled.
 
@@
 \srccm{}s memory is split in three: heap, free stack and encapsulated local stack frames.
-The three kinds of memory and their relative position to each other are sketched in Figure~\ref{fig:mem-sketch}.
-Note how the encapsulated local stack frames are sparsely laid out in the stack with free stack in between.
-The free stack in between stack frames may be used by untrusted code for their local stack frames.
-Untrusted stack frames are not protected by \srccm{}, but this does not prevent untrusted code from securing their own stack frames.
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-
-  \scope
-    \clip (-.1,-.1) rectangle (4.6,11.1);
-    \fill[fill=white] (0,0) rectangle (4.5,11);
-    \draw (0,0) -- (0,11);
-    \draw (4.5,0) -- (4.5,11);
-  \endscope
-  \draw (-1,0) node {};
-  \draw (-1,12) node {};
-
-  \draw[pattern=dots,draw=none] (0,0) rectangle (4.5,2);
-  \draw[pattern=dots,draw=none] (0,9.5) rectangle (4.5,11);
-
-  \draw[pattern=north west lines] (0,2) rectangle (4.5,4);
-  \draw[pattern=crosshatch] (0,4) rectangle (4.5,5);
-  \draw[pattern=north west lines] (0,5) rectangle (4.5,6);
-  \draw[pattern=crosshatch] (0,6) rectangle (4.5,7.5);
-  \draw[pattern=north west lines] (0,7.5) rectangle (4.5,8.5);
-  \draw[pattern=crosshatch] (0,8.5) rectangle (4.5,9.5);
-
-  \draw (5.75,2.25) node[anchor=west] {\tiny $\stkb$};
-  \draw (5.75,2.25) edge[-latex] (4.75,2.25);
-  \draw (5,9.5) edge[-latex] (5,8);
-  \draw [decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt},yshift=0pt] (-0.25,9.5) -- (-0.25,2) node [black,midway,xshift=-.75cm,yshift=0.05cm,rotate=45] {\footnotesize stack};
-
-  \matrix [draw,below right] at (current bounding box.north east) {
-    \node[pattern=dots,draw,label=right:{\tiny Heap ($\ms$)}] {}; \\
-    \node[pattern=north west lines,draw,label=right:{\tiny Free stack ($\ms_\stk$)}] {}; \\
-    \node[pattern=crosshatch,draw,label=right:{\tiny Encapsulated stack frame ($\stk$)}] {}; \\
-};
-
-  \end{tikzpicture}
-  \caption{A sketch of how heap, encapsulated stack and free stack are laid out in memory.
-    The encapsulated stack frames and the free stack constitutes the stack.
-    The encapsulated stack frames may have free stack in between them where stack frames of non-trusted code may reside.
-    The stack grows downwards in memory with $\stkb$ as the top address (and the base address of the free stack capability).}
-  \label{fig:mem-sketch}
-\end{figure}
-
 In order to model the three kinds of memory, we simply have three sub-worlds.
-That is, our world is defined as
+That is, our world is defined as a product:
 \[
@@
 In order to define what this actually means, we need to be more precise about what we mean by invariants.
-We call the invariants regions because, as we will see later, they turn out to not be invariant.
+
+We call the invariants regions because, as we will see in Section~\ref{subsubsec:ft-and-revocation}, they turn out to not be invariant.
 A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
@@
 In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some the intuition.
-Say we have two integers, they are related if they are equal.
+If we have two integers, then they are related when they are equal.
 If we instead have two capabilities, then they should also be related if they, in some sense, are equal.
-But what does it mean for capabilities to be ``equal''?
+But what does it mean for capabilities to be \textit{equal}?
 Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably perform the same computation.
@@
 % Linearity
-The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they can reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
+The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
 In order to model this uniqueness, we need to keep track of which parts of memory that are uniquely referenced and make sure that only one linear capability references the unique parts.
@@
 We use the world to keep track of what parts of memory must be uniquely referenced by having two kinds of regions: shared and spatial.
-If a memory segment is governed by a shared region, then non-linear capabilities may reference it.
+If a memory segment is governed by a shared region, then normal capabilities may reference it.
 On the other hand, if a memory segment is governed by a spatial region, then only a linear capability may reference it.
@@
 We cannot let multiple linear capabilities reference the same memory, so we add ownership to spatial regions.
-It is only the spatial owned regions that can be referenced by a capability.
+Only the spatial owned regions can be referenced by a capability.
 Even though only spatial owned regions can be referenced by linear capabilities, we still have unclaimed spatial regions that specify what part of memory looks like.
-When there is an unclaimed spatial region, then there may be a compatible world  that claims the ownership of that spatial region which means that the memory can be referenced (we return to the notion of compatible worlds w.r.t.
-ownership later).
+When there is an unclaimed spatial region, then there may be a compatible world  that claims the ownership of that spatial region which means that the memory may still be referenced (we return to the notion of compatible worlds w.r.t.
+ownership in Section~\ref{subsubsec:joining-worlds}).
 Specifically, we add tags $\spatial$ and $\spatialo$ to the spatial regions:
@@
 \]
-% TODO is the below notation even used?
-For readability we will sometimes use $\spao$ as short for $\spatialo$, $\spa$ as short for $\spatial$, and $\pur$ as short for $\pure$.
 
 % What parts of memory are linear
-We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we have introduced enough to finish the definition of each of the three sub-worlds.
-The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and non-linear capabilities, so it should contain both shared and spatial regions.
+We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we continue the definitions of the three sub-worlds.
+The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and normal capabilities, so it should contain both shared and spatial regions.
 For this reason, it is defined as
@@
 Given a world, we want to be able to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
-This is expressed by stripping the world of all its ownership, so the only active regions are the $\pure$ regions.
+This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions.
 We refer to this as the pure part of the world and define the function $\purePart{}$ which removes all ownership.
@@
 The seal interpretation function should be an invariant, so we need to add it to a type of region that is invariant.
-As we will see in Section~\ref{subsubsec:ft-and-revocation}, shared regions are invariant, so we add the seal interpretation function as follows
+Shared regions are invariant, so we add the seal interpretation function to it:
 \begin{multline*}
@@
 % Memory evolves over time, we need to model this.
-The world specifies the contents of the memory.
-When a machine executes, the contents of the memory changes and parts of memory may be repurposed.
-The world must model these changes.
+During machine execution, the memory contents changes and parts of memory may be repurposed.
+The world must model such memory changes.
 % Future World
-To this end, we have a future world relation $\future$ that describes how worlds may change over time.
-For our purpose, relating \trgcm{} to \srccm{}, we are mainly interested in changes to the stack as it is the part that will be repurposed.
+To this end, we have a future world relation $\future$ that describes how worlds and thus memory may change over time.
+In order to relate \trgcm{} and \srccm{}, we are mainly interested in changes to the stack as it repurposes stack frames.
 Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions that govern call frames that have been popped from the call stack, so the stack frame can be reused for a new call frame.
@@
 \lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).}
-For technical reasons, we need the worlds to be extensional which means that when it comes to removing regions in future worlds, we cannot simply forget about them.
-Instead of removing the regions, we keep them around and tag them as removed with a $\revoked$ tag.
+For technical reasons, we need the worlds to be extensional, so we cannot simply forget about regions that we want to remove.
+We keep removed regions in the world and tag them as removed with a $\revoked$ tag.
 This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.
-It is not all regions that need to be able to be revoked.
-Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities in some sense can be revoked.
-A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory.
-This means that it is safe to repurpose that part of memory.
-On the other hand, non-linear capabilities may have an alias, so it is not safe to repurpose the memory a non-linear capability has authority over.
-It is only spatial regions that specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
+Not all regions that should be revocable.
+Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities, in some sense, can be revoked.
+A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory which makes it safe to repurpose.
+On the other hand, normal capabilities may have an alias, so it is not safe to repurpose the memory a normal capability has authority over.
+Only spatial regions specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
 The $\revoked$ region does not specify any memory, so we drop the world indexed relation.
@@
 This means that a pair of memory segments that are related now will stay related in future worlds.
-Our spatial regions now look like this:
+This changes the spatial regions as follows:
 \[
@@
 \]
-We make a similar change for $\Regionh$.
+We make a similar change to $\Regionh$.
 % TODO update terminology: memory invariant, H, and seal invariant, H_\sigma
@@
 \subsubsection{Joining worlds}
-% TODO rewrite this introduction, add text between definitions (?), and reread this section.
+\label{subsubsec:joining-worlds}
 The world serves multiple purposes as it is both a specification of memory contents as well as a specification of authority.
 This is best seen in the operators used to join worlds.
-First when we see the world as a memory specification, we have a pretty standard join $\uplus$ that simply requires that the two worlds do not use the same region names.
-\begin{definition}[World disjoint union $\uplus$]
+First when we see the world as a memory specification, we have a pretty standard join $\uplus$ that simply requires the worlds to have different region names.\begin{definition}[World disjoint union $\uplus$]
   Given worlds $W_1$, $W_2$, $W$
@@
 
-% TODO add figure that illustrates this.
 According to \stktokens{} (described in \sectionname~\ref{sec:stktokens-explained}), stack frames must be non-empty, so they are distinguishable from a missing stack frame.
@@
 This means that the protected stack frames are not necessarily packed tightly in memory, and the memory in between is part of the free stack.
-% TODO illustrate the protected stack ?
+The free stack in between stack frames may be used by untrusted code for their local stack frames (Untrusted stack frames are not protected by \srccm{}, but this does not prevent untrusted code from securing their own stack frames).
+\figurename~\ref{fig:mem-sketch} sketches this.
+
 Each stack frame in the \srccm{} stack contains an old program pointer which corresponds to the old program counter recorded in the region of the world associated with the stack frame.
 To achieve this, the partition function $R_\ms$ also records an $\opc$ for each region, and this $\opc$ should establish the link between the region and the stack frame.
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+
+  \scope
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+  \endscope
+  \draw (-1,0) node {};
+  \draw (-1,12) node {};
+
+  \draw[pattern=north west lines,draw=none] (0,0) rectangle (4.5,2);
+  \draw[pattern=north west lines,draw=none] (0,9.5) rectangle (4.5,11);
+
+  \draw[pattern=none] (0,2) rectangle (4.5,4);
+  \draw[pattern=bricks] (0,4) rectangle (4.5,5);
+  \draw[pattern=none] (0,5) rectangle (4.5,6);
+  \draw[pattern=bricks] (0,6) rectangle (4.5,7.5);
+  \draw[pattern=none] (0,7.5) rectangle (4.5,8.5);
+  \draw[pattern=bricks] (0,8.5) rectangle (4.5,9.5);
+
+  \draw (5.75,2.25) node[anchor=west] {\footnotesize $\stkb$};
+  \draw (5.75,2.25) edge[-latex] (4.75,2.25);
+  \draw (5,9.5) edge[-latex] (5,8);
+%  \draw [decorate,decoration={brace,amplitude=7pt,mirror,raise=4pt},yshift=0pt] (-0.25,9.5) -- (-0.25,2) node [black,midway,xshift=-1cm,yshift=0.05cm,rotate=45] {stack};
+  \draw [-|] (-0.5,9.5) edge (-0.5,2);
+  \draw (-0.75,5.75) node[anchor=east] [black] {stack};
+  \matrix [draw,below right] at (current bounding box.north east) {
+    \node[pattern=north west lines,minimum size=0.4cm,draw,label=right:{\footnotesize Heap ($\ms$)}] {}; \\
+    \node[minimum size=0.4cm,draw,label=right:{\footnotesize Free stack ($\ms_\stk$)}] {}; \\
+    \node[pattern=bricks,minimum size=0.4cm,draw,label=right:{\footnotesize Encapsulated stack frame ($\stk$)}] {}; \\
+};
+
+  \end{tikzpicture}
+  \caption{A sketch of how heap, encapsulated stack and free stack are laid out in memory.
+    The encapsulated stack frames and the free stack constitutes the stack.
+    The encapsulated stack frames may have free stack in between them where stack frames of non-trusted code may reside.
+    The stack grows downwards in memory with $\stkb$ as the top address (and the base address of the free stack capability).}
+  \label{fig:mem-sketch}
+\end{figure}
 
@@
 Generally speaking, an observation relation captures the property we want to prove.
-% TODO double check "co-termination" (equitermination?)
 Ultimately, we want to prove a full-abstraction theorem which is defined in terms of contextual equivalence for components that in turn is defined as co-termination in any context.
@@
 For this reason, we define two value relations a trusted $\lrvtrusted$ and an untrusted $\lrv$.
-It is safe to give trusted programs anything that is safe for an untrusted program, so the trusted value relation is defined as a super set of the untrusted value relation.
+Anything safe for unstrusted programs is also safe to give to a trusted program, so the trusted value relation is defined as a super set of the untrusted value relation.
 
@@
         \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \\
-        \quad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
+        \quad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spatialo,\square}, \aaddr'+\calllen) \tand \\
         \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
@@
 The code capability for a closure must point to the code memory because it is the only part of memory that is executable.
-It is not safe for untrusted components to have a capability for a trusted components code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
+Untrusted components cannot safely have a capability for a trusted components code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
 While it is not safe to give a bare capability for a trusted components code memory, it can be perfectly safe to give a sealed capability for a trusted components code.
@@
 
-When it comes to the data capability of a closure, we just require that it comes from the untrusted value relation because the trusted value relation contains nothing that makes sense to seal as the data capability
-% TODO be more precise, what later paragraph means.
-(we return to the specific contents of the two value relations in a later paragraph).
+When it comes to the data capability of a closure, we just require that it comes from the untrusted value relation because the trusted value relation contains nothing that makes sense to seal as the data capability (we return to the specific contents of the two value relations later in this section).
 
@@
 \]
-The code region is $\pure$ because it needs to contain a seal invatiant and because we assume that code pointers are non-linear.
-
-% TODO in the above, add something about the untrusted case where the seals should nut be in the "global" seals because the global seals are the one that we globally agree to be trusted.
+The code region is $\pure$ because it needs to contain a seal invatiant and because we assume that code pointers are normal capabilities.
+
+% TODO in the above, add something about the untrusted case where the seals should not be in the "global" seals because the global seals are the one that we globally agree to be trusted.
 
@@
 % TODO add brief intro to paragraph
-
 The world can be seen as an authority specification which means that it dictates what kind of capabilities can address a certain part of memory.
-Specifically, linear capabilities can only address memory governed by a $\spatialo$ region, and non-linear capabilities can only address memory governed by a $\pure$ region.
+Specifically, linear capabilities can only address memory governed by a $\spatialo$ region, and normal capabilities can only address memory governed by a $\pure$ region.
 All the permission based condition we define project the regions that the capability may address from the world.
@@
 If we need to establish the $\readCond{}$ after a splice, we can simply use the union of the regions that witnessed the $\readCond{}$ of the two individual capabilities.
-We also need to support splitting which is no problem for non-linear capabilities as the same $\pure$ region can be used to establish the $\readCond{}$ for multiple non-linear capabilities.
+We also need to support splitting which is no problem for normal capabilities as the same $\pure$ region can be used to establish the $\readCond{}$ for multiple normal capabilities.
 On the other hand, a $\spatialo$ region can only be used to establish the $\readCond{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
@@
 When we define the logical relation, we even require the executable capability to not be linear.
-It is not clear whether a linear executable capability would even be useful because it cannot be moved from the $\pcreg$-register without crashing the execution.
+Linear executable capabilities would likely not be useful because they cannot be moved from the $\pcreg$-register without crashing the execution.
 This may sound like an ideal primitive for constructing something that can be executed once, however, most programs rely on loading other capabilities or seal sets using the program counter capability which is not possible when the program counter is linear.
@@
 The first case is data.
-It is always safe to posses data as data is just integers that grant no authority.
+Data grant no authority, so data is always safe.
 Further unlike capabilities, it is always possible to construct a new integer with the move instruction.
@@
 That is, if a capability has read permission, then $\readCond{}$ must be satisfied, if it has write permission, then $\writeCond{}$ must be satisfied, and if it has execute permission, then $\execCond{}$ and $\xReadCond{}$ must be satisfied.
-If the capability has execute permission, then it must also be a non-linear capability.
+If the capability has execute permission, then it must also be a normal capability.
 Finally, the capability cannot have read/write/execute permission because that would break the write-XOR-execute assumption, i.e.\ the code memory in non-writable and data memory is non-executable.
@@
 % Code pointers for trusted code
-Two capabilities are related as trusted code pointers if they are non-linear, has a permission derivable from read-execute, and has a range of authority within the trusted address space, $\ta$.
+Two capabilities are related as trusted code pointers if they are normal, has a permission derivable from read-execute, and has a range of authority within the trusted address space, $\ta$.
 Further, we need to know that the capabilities actually point to a piece of code which is why we require the $\xReadCond{}$ to be satisfied.
@@
 The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrv[]$, so the read condition would not hold.
-It is, however, fine to let trusted code have access to such seals because we expect the trusted code to treat the seals reasonably.
+However, trusted code can have access to such seals because we expect the trusted code to treat the seals reasonably.
 \begin{figure}

### 2019-07-17 popl-2019-journal-v/paper.tex
[POPLjv] Add diplay math and remove todo
Add a paragraph at the end of the introduction about the relation to the POPL paper.
Needs further substantiating.

@@
-% TODO do we say something about linear pcs and why they don't make sense? (related to def:reasonable-pc)
-
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
@@
 \end{itemize}
-This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.
+
+\begin{jversion}
+  This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens:_2019}.
+  Compared to the earlier text, this version adds X and Y and Z.
+  \todo{new version of TR?}
+  This paper is accompanied by a technical report~\citep{technical_report_popl} with the elided details and proof.
+\end{jversion}
 
@@
     We say that a component
-    \[(\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\]
+    \[
+      (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+    \]
     is reasonable if the following hold: For all $(s \mapsto \sealed{\sigma,\vsc}) \in

### 2019-07-17 popl-2019-journal-v/paper.tex
Some minor edits

@@
   \subsection{The purpose of sealing}
+  \label{sec:purpose-sealing}
   % Sealing example:
@@
 Rather than specifying a decode function and an encode function, we assume that they are given with certain properties.
-The $\decInstr{} : \Word \fun \Instr$ should be injective and surjective for all non-$\fail$ instructions.
+The $\decInstr{} : \Word \fun \Instr$ should be surjective and injective for all non-$\fail$ instructions.
 Further, it should decode all capabilities as the $\fail$ instruction, i.e.\
@@
 In Section~\ref{subsec:overlay-semantics}, we present an overlay semantics which changes the set of words compared to this machine.
+\todo[inline]{Dominique: perhaps delay the discussion of encodetype for the overlay semantics to Section~\ref{subsec:overlay-semantics}?}
 The overlay semantics also has a type encoding function which must encode the new words, so they correspond to their representation on this machine.
@@
 When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
-Return pointers, on the other hand, are replaced by a pair of sealed return capabilities.
+Return pointers, on the other hand, are replaced by a pair of sealed return capabilities, as we already explained in Section~\ref{sec:purpose-sealing}.
 % KJAA: Again with the sealed pair. Again, this idea is unclear.
@@
 % Prevention of attack 2
-In a traditional calling convention with a single stack, the stack serves as a call stack keeping track of the order calls where made in and thus in which order they should be returned to.
+In a traditional calling convention with a single stack, the stack serves as a call stack keeping track of the order calls were made in and thus in which order they should be returned to.
 A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
@@
 The former is sealed and used for the data part of the return pair.
-\item \label{item:return-code} Create code return capability on call: Seal the old program pointer.
-\item Reasonable use of seals: Return seals are only used to seal old program pointers, every return seal is only used for one call site, and they are not leaked.
+\item \label{item:return-code} Create code return capability on call: Seal the old program counter capability.
+\item Reasonable use of seals: Return seals are only used to seal old program counter capabilities, every return seal is only used for one call site, and they are not leaked.
 \end{enumerate}
@@
   \centering
-{\footnotesize
+{\small
 \[

### 2019-07-17 popl-2019-journal-v/paper.tex
Some rephrasing and a suggestion.

@@
 % - Address must be in data memory
+\todo[inline]{Dominique: a bit of intro seems needed here: explain that we require certain syntactic properties corresponding to a linker doing its job properly and this is captured by a a syntactic well-formedness judgement that we explain in this section.}
+
 \begin{jversion}
@@
   All the available seals of a component should reside in the code memory.
-  Both adversarial and trusted components can be well-formed.
-  Adversarial components are components where the range of addresses for the code memory is disjoint from the trusted address space Further, adversarial code does not have access to return seals.
-  On the other hand, trusted components must make return seals available to all its calls in a way that is consistent with \stktokens{}.
+
+  As discussed before, we only provide guarantees (LSE and WBCF) for components respecting certain semantic rules (reasonable use of seals).
+  Components that do not satisfy these requirements are still allowed in \srccm{}, but (as we will see below) function calls in such components will not behave according to the special \srccm{} semantics that guarantees LSE and WBCF.
+  To distinguish trusted components (which satisfy reasonability requirements) from adversarial components (which don't), we make use of a set of trusted addresses $\ta$.
+  Any component's code memory must fall entirely within or outside of $\ta$, so every component is either trusted or adversarial.
+  Some well-formedness requirements are specific for adversarial and trusted components.
+  Well-formed adversarial components cannot have return seals.
+  On the other hand, well-formed trusted components must make return seals available to all its calls in a way that is consistent with \stktokens{}.
   This means that each return seal is only used for one call.
-  Finally, well-formed components can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
+
+  Finally, well-formed components (adversarial and trusted) can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
 \end{jversion}
@@
 \begin{jversion}
-The judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
+Formally, the judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
 The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.

### 2019-07-17 popl-2019-journal-v/paper.tex
Move figures closer to where they're referenced.

@@
   Well-formed adversarial components cannot have return seals.
-  On the other hand, well-formed trusted components must make return seals available to all its calls in a way that is consistent with \stktokens{}.
+  On the other hand, well-formed trusted components can have return seals, but must make return seals available to all its calls in a way that is consistent with \stktokens{}.
   This means that each return seal is only used for one call.
 
+  \todo[inline]{Dominique: The following line appears to explain an arbitrary point about well-formedness: I would explain here that well-formedness imposes a quite rigid structure on well-formed components: a component's code memory may contain only data and sets of seals, while the data memory contains only data, non-sealed memory capabilities to the component's data or code memory (respecting linearity) or sealed memory or code capabilities (to the respective part of the component's memory, sealed with closure seals).
+A component's exports should be sealed code or data capabilities, sealed with a closure seal.
+    Also: perhaps move this before the point about adversarial versus trusted components.
+  }
   Finally, well-formed components (adversarial and trusted) can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
@@
 % The exports $\overline{\var{export}}$ can be anything non-linear allowed to reside in data memory or a sealed capability for the code memory sealed with one of the closure seals.
+
+\begin{figure}
+\begin{mathpar}
+  \inferrule*[Right=Base]{
+    \dom(\mscode) = [\baddr,\eaddr]\\
+    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
+    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
+    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}\\
+    \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
+    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
+    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} \\
+    \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}} \\
+    \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
+    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} \\
+    (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset) \\
+    \dom(\msdata) \mathrel{\#} \ta \\
+    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
+    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
+    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
+  }{
+    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
+  }
+  \and
+  \inferrule*[Right=Main]{
+    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
+    \ta \vdash \var{comp}_0 \\
+    (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
+  }{
+    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
+  }
+\end{mathpar}
+\caption{Well-formedness judgement.}
+\label{fig:well-formed}
+\end{figure}
 
@@
 
+\begin{figure}
+  \begin{subfigure}[h]{1.0\linewidth}
+    \begin{mathpar}
+  \inferrule*[Right=C-Seals]{
+    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \\ [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
+  }{
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
+  }
+  \and
+  \inferrule*[Right=C-Instr]{
+    \mscode(a) \in \ints\\
+{    \begin{multlined}
+      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
+      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
+    \end{multlined}
+}  }{
+    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
+  }
+  \and
+  \inferrule*[Right=C-Mem]{
+    \text{$\mscode$ has no hidden calls}\\
+    \sigrets \mathrel{\#} \sigcloss \\
+    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
+    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
+    \forall a \in \dom(\mscode) \ldotp 
+    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
+    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
+  }{
+    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
+  }
+    \end{mathpar}
+    \caption{Code well-formedness.}
+    \label{fig:well-formed-code}
+  \end{subfigure}
+
+  \begin{subfigure}[h]{1.0\linewidth}
+    \begin{mathpar}
+      \inferrule*[Right=W-Data]{
+        z \in \ints
+  }{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} z
+  }
+  \and
+  \inferrule*[Right=W-Capability]{
+    \permbnf \sqsubseteq \rw\\
+    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}} \\\\
+    \lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
+  }{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
+  }
+  \and
+  \inferrule*[Right=W-Sealed-Capability]{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \vsc \\
+    \sigma \in \sigcloss
+  }{
+    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \sealed{\sigma,\vsc}
+  }
+    \end{mathpar}
+    \caption{Word well-formedness.}
+    \label{fig:well-formed-word}
+  \end{subfigure}
+
+  \begin{subfigure}[h]{1.0\linewidth}
+    \begin{mathpar}
+  \inferrule*[Right=E-Sealed-Code]{
+    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
+    \sigma \in \sigcloss
+  }{
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
+  }
+  \and
+  \inferrule*[Right=E-Word]{
+    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
+  }{
+    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
+  }
+    \end{mathpar}
+    \caption{Export well-formedness}
+    \label{fig:well-formed-export}
+  \end{subfigure}
+  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
+  \label{fig:well-formed-aux}
+\end{figure}
 \figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ which specifies the words that are well-formed in a component's data memory.
@@
 
-\begin{figure}
-\begin{mathpar}
-  \inferrule*[Right=Base]{
-    \dom(\mscode) = [\baddr,\eaddr]\\
-    [\baddr-1,\eaddr+1] \mathrel{\#} \dom(\msdata)\\
-    \mspad = [\baddr-1\mapsto 0] \uplus [\eaddr+1 \mapsto 0]\\
-    \exists A_\mathrm{own} : \dom(\msdata) \rightarrow \powerset{\dom(\msdata)}\\
-    \dom(\msdata) = A_{\mathrm{non-linear}} \uplus A_\linear \\
-    A_\linear = \biguplus_{a \in \dom(\msdata)} A_{\mathrm{own}}(a) \\
-    \overline{\var{export}} = \overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}} \\
-    \overline{\var{import}} = \overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}} \\
-    \{\overline{a_{\mathrm{import}}}\} \subseteq \dom(\msdata)\\
-    \overline{s_{\mathrm{import}}} \mathrel{\#} \overline{s_{\mathrm{export}}} \\
-    (\emptyset \neq \dom(\mscode) \subseteq \ta) \vee (\dom(\mscode) \mathrel{\#} \ta \wedge \sigrets = \emptyset) \\
-    \dom(\msdata) \mathrel{\#} \ta \\
-    \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode \\
-    \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
-    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
-  }{
-    \ta \vdash (\mscode\uplus \mspad,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)
-  }
-  \and
-  \inferrule*[Right=Main]{
-    \var{comp}_0 = (\mscode,\msdata,\overline{\var{import}},\overline{\var{export}},\sigrets,\sigcloss,A_\linear)\\
-    \ta \vdash \var{comp}_0 \\
-    (\_ \mapsto c_{\mathrm{main},c}), (\_ \mapsto c_{\mathrm{main},d}) \in \overline{\var{export}}
-  }{
-    \ta \vdash (\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})
-  }
-\end{mathpar}
-\caption{Well-formedness judgement.}
-\label{fig:well-formed}
-\end{figure}
-
-\begin{figure}[h]
-  \begin{subfigure}[h]{1.0\linewidth}
-    \begin{mathpar}
-  \inferrule*[Right=C-Seals]{
-    \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \\ [\sigma_\baddr,\sigma_\eaddr] = (\sigrets \cup \sigcloss)
-  }{
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
-  }
-  \and
-  \inferrule*[Right=C-Instr]{
-    \mscode(a) \in \ints\\
-{    \begin{multlined}
-      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
-      \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
-    \end{multlined}
-}  }{
-    \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
-  }
-  \and
-  \inferrule*[Right=C-Mem]{
-    \text{$\mscode$ has no hidden calls}\\
-    \sigrets \mathrel{\#} \sigcloss \\
-    \exists d_\sigma : \dom(\mscode) \rightarrow \powerset{\Seal} \ldotp
-    \sigrets = \biguplus_{a \in \dom(\mscode)} d_\sigma(a) \tand\\
-    \forall a \in \dom(\mscode) \ldotp 
-    \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
-    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
-  }{
-    \sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode
-  }
-    \end{mathpar}
-    \caption{Code well-formedness.}
-    \label{fig:well-formed-code}
-  \end{subfigure}
-
-  \begin{subfigure}[h]{1.0\linewidth}
-    \begin{mathpar}
-      \inferrule*[Right=W-Data]{
-        z \in \ints
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} z
-  }
-  \and
-  \inferrule*[Right=W-Capability]{
-    \permbnf \sqsubseteq \rw\\
-    \lin = \linear \Rightarrow \emptyset \subset [\baddr,\eaddr] \subseteq A_{\mathrm{own}} \\\\
-    \lin = \normal \Rightarrow [\baddr,\eaddr] \subseteq A_{\mathrm{non-linear}}
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} ((\perm,\lin),\baddr,\eaddr,\aaddr)
-  }
-  \and
-  \inferrule*[Right=W-Sealed-Capability]{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \vsc \\
-    \sigma \in \sigcloss
-  }{
-    A_{\mathrm{code}},A_{\mathrm{own}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \sealed{\sigma,\vsc}
-  }
-    \end{mathpar}
-    \caption{Word well-formedness.}
-    \label{fig:well-formed-word}
-  \end{subfigure}
-
-  \begin{subfigure}[h]{1.0\linewidth}
-    \begin{mathpar}
-  \inferrule*[Right=E-Sealed-Code]{
-    [\baddr,\eaddr] \subseteq A_{\mathrm{code}} \\
-    \sigma \in \sigcloss
-  }{
-    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto \sealed{\sigma,((\rx,\normal),\baddr,\eaddr,\aaddr)}
-  }
-  \and
-  \inferrule*[Right=E-Word]{
-    A_{\mathrm{code}},\emptyset,A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} w
-  }{
-    A_{\mathrm{code}},A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-export}} s \mapsto w
-  }
-    \end{mathpar}
-    \caption{Export well-formedness}
-    \label{fig:well-formed-export}
-  \end{subfigure}
-  \caption{Well-formedness judgement for code, words, and export. $\calllen$ it the length of the call code.}
-  \label{fig:well-formed-aux}
-\end{figure}
 \end{jversion}

### 2019-07-17 popl-2019-journal-v/paper.tex
[POPLjv] Update references

@@
 The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
-More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
+More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri_2015}.
 These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
@@
 \begin{jversion}
-  This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens:_2019}.
+  This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens_2019}.
   Compared to the earlier text, this version adds X and Y and Z.
@@
 In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri_2015} with linear capabilities as the main addition.
 For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
@@
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015,watson_fast_2016}.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri_2015,watson_fast_2016}.
 % seals
@@
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for details.
+Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.
 \begin{figure}[tb]
@@
 
-Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
-The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
+Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri_2015} for an overview of previous work.
+The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri_2015,woodruff_cheri_2014}.
 This is a recent, relatively mature capability machine which combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
@@
 
-CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
+CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri_2015}.
 The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.

### 2019-07-17 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Work on the proving full abstraction section

@@
 It also splits the world in three to distribute the authority of the world.
-The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrheap$, $\lrstk[]$, and $\lrfree[]$, respectively. 
-The $\lrheap$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
-The $\lrheap$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
+The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrhname$, $\lrstk[]$, and $\lrfree[]$, respectively. 
+The $\lrhname$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
+The $\lrhname$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
 The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
@@
 \label{def:heap-rel}
-For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrheap$ as:
+For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrhname$ as:
 \[
@@
 Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
-The $\lrheap$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.
-
-To a large extend, the definition of $\lrheap$ is pretty standard.
-$\lrheap$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
+The $\lrhname$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.
+
+To a large extend, the definition of $\lrhname$ is pretty standard.
+$\lrhname$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
 The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction partitions the authority of the world using $\oplus$.
@@
 The heap sub-world contains all seal invariants.
-Similar to memory segments, only one seal invariant should impose restrictions on a seal, which $\lrheap$ makes sure is the case.
+Similar to memory segments, only one seal invariant should impose restrictions on a seal, which $\lrhname$ makes sure is the case.
 \begin{definition}[Free stack relation]
@@
 \end{definition}
-The free stack relation $\lrfree[]$ is in most regards like the heap relation, $\lrheap{}$.
+The free stack relation $\lrfree[]$ is in most regards like the heap relation, $\lrhname{}$.
 The free stack relation, partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under part of the world.
@@
 A value relation based on what it is safe for an adversary to have should prohibit return seals, so such a relation cannot be used to reason about trusted programs.
-For this reason, we define two value relations a trusted $\lrvtrusted$ and an untrusted $\lrv$.
+For this reason, we define two value relations a trusted $\lrvname_{\trusted}$ and an untrusted $\lrvname$.
 Anything safe for unstrusted programs is also safe to give to a trusted program, so the trusted value relation is defined as a super set of the untrusted value relation.
@@
 The untrusted and trusted value relation both relates \srccm{} and \trgcm{} words.
-The untrusted value relation $\lrv$ relates words that are safe to give to untrusted programs and $\lrvtrusted$ relates words that are safe to give to trusted programs.}
+The untrusted value relation $\lrvname$ relates words that are safe to give to untrusted programs and $\lrvname_\trusted$ relates words that are safe to give to trusted programs.}
 \label{fig:value-relation}
@@
 The $\execCond{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
-We later define the $\lre$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
+We later define the $\lrename$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
 An executable capability can have its range of authority shrunk or its current address changed which changes what instructions are executed and thus potentially whether the code respects memory invariants.
@@
 \label{par:untrusted-val-rel}
-The untrusted value relation $\lrv$ relates all the words that untrusted components can safely posses.
+The untrusted value relation $\lrvname$ relates all the words that untrusted components can safely posses.
 That is words that cannot be used to break any memory invariants.
@@
 Further, if we take any sealed capabilities related under $\sigma$ in the seal invariant, then they should produce a safe result.
-We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrexj$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
+We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrename_\mathrm{xjmp}$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
 % Linear and non-linear case. Affects the future world we consider.
@@
 \label{par:trusted-val-rel}
-% TODO write section
-The trusted value relation $\lrvtrusted$ relates everything safe for a trusted component to have without breaking memory invariants.
+The trusted value relation $\lrvname_\trusted$ relates everything safe for a trusted component to have without breaking memory invariants.
 For the most part, we allow them to contain the same words as the untrusted components, but we also need to allow them to have seal sets with trusted closure seals and return seals which we cannot allow untrusted components to have.
@@
 Note that even though the capability has read permission, we do not require the read condition to hold.
-The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrv[]$, so the read condition would not hold.
+The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrvname$, so the read condition would not hold.
 However, trusted code can have access to such seals because we expect the trusted code to treat the seals reasonably.
@@
 
-\caption{The trusted value relation $\lrvtrusted$ relates all the words that are safe for trusted components to contain.
+\caption{The trusted value relation $\lrvname_\trusted$ relates all the words that are safe for trusted components to contain.
 A trusted component may contain untrusted words (\figurename~\ref{fig:untrusted-val-rel}), return seals and trusted closure seals, and code pointers for trusted code.
@@
 We use this to not relate register content that will be overwritten anyway.
-We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$.
+We write $\lrrname(W)$ to mean $\lrrname(\emptyset)(W)$.
 That is, if we do not need to exclude additional registers, then we simply omit that argument.
@@
 \subsubsection{Expression relations}
-% TODO write section
-The expression relation $\lre$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e.
+The expression relation $\lrename$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e.
 the capabilities can be used to construct configurations in the observation relation.
-The $\lre$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control flow during execution when a $\tjmp{}$ instruction is executed.
+The $\lrename$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control flow during execution when a $\tjmp{}$ instruction is executed.
 In the setting of \srccm{} and \trgcm{}, we can also use sealed capabilities to change the control flow by using the $\txjmp{}{}$ instruction.
-The $\txjmp{}{}$ instruction updates the $\pcreg$ register and the $\rdata$ register, however, the $\lre$ relation only updates the $\pc$-register, so we cannot use $\lre$ to reason about sealed capabilities.
-Instead, we define the relation $\lrexj$ which relates two pairs of capabilities when they are safe to use with the $\txjmp{}{}$ instruction.
+The $\txjmp{}{}$ instruction updates the $\pcreg$ register and the $\rdata$ register, however, the $\lrename$ relation only updates the $\pc$-register, so we cannot use $\lrename$ to reason about sealed capabilities.
+Instead, we define the relation $\lrename_\mathrm{xjmp}$ which relates two pairs of capabilities when they are safe to use with the $\txjmp{}{}$ instruction.
 
@@
 Instead, we need to pick configurations made out of related components, i.e.\ related register files and related memories that respect linearity.
-The type of execution captured by $\lre$ corresponds to a normal jump.
+The type of execution captured by $\lrename$ corresponds to a normal jump.
 When a $\tjmp r$ instruction is interpreted, the $\pcreg$-register is replaced with the contents of register $r$, i.e. the current configuration is plugged with a new pc.
-The $\lre$ relation is defined in \figurename~\ref{fig:expr-rels}
-
-The $\lrexj$ relation looks very much like the $\lre$ relation.
+The $\lrename$ relation is defined in \figurename~\ref{fig:expr-rels}
+
+The $\lrename_\mathrm{xjmp}$ relation looks very much like the $\lrename$ relation.
 It takes related memories and register files (ignoring the $\rdata$ register) and combines them into two configurations.
-Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lro$ relation.
-The $\lrexj$ relation is defined in \figurename~\ref{fig:expr-rels}.
+Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lroname$ relation.
+The $\lrename_\mathrm{xjmp}$ relation is defined in \figurename~\ref{fig:expr-rels}.
 \begin{figure}
@@
 This means that we need to lift relatedness one step further than the components, namely to the level of execution configurations.
-To this end, we define $\lrec$.
+To this end, we define $\lrecname$.
 \begin{definition}[Related execution configuration]
@@
 Definition~\ref{def:exec-conf-rel} essentially says that two executable configurations are related when they are made out of related components.
-That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$ and $\rdata$ registers are related by the $\lrexj$ relation.
+That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$ and $\rdata$ registers are related by the $\lrename_\mathrm{xjmp}$ relation.
 Further, the two memories and the two register files should be related.

### 2019-07-18 popl-2019-journal-v/appendix.tex
[POPLjv] Update split and splice in appendix to work on seal

@@
     \hline
-   %TODO change to seal
-    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_3}{w}\\ &\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $\Phi(\rn) = n \in \nats$ and $\baddr \le n < \eaddr$ and $c_1 = ((\perm,\lin),\baddr,n,\aaddr)$ and $c_2 = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $w = \linCons{\Phi(r_3)}$\\
+    \tsplit{r_1}{r_2}{r_3}{\rn}                  & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{c_1}\\ &\updReg{r_2}{c_2})\endarray & $\Phi(r_3) = \seal{\sigma_b,\sigma_e,\sigma}$ and $\Phi(\rn) = \sigma_n \in \nats$ and $\sigma_b \le \sigma_n < \sigma_e$ and $c_1 = \seal{\sigma_b,\sigma_n,\sigma}$ and $c_2 = \seal{\sigma_n+1,\sigma_e,\sigma}$\\
     \hline
-   % TODO seal
-    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray& $\Phi(r_2) = ((\perm,\lin),\baddr,n,\_)$ and $\Phi(r_3) = ((\perm,\lin),n+1,\eaddr,\aaddr)$ and $\baddr \le n < \eaddr$ and $c = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$\\
+    \splice{r_1}{r_2}{r_3}                       & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{c})\endarray& $\Phi(r_2) = \seal{\sigma_b,\sigma_n,\_}$ and $\Phi(r_3) = \seal{\sigma_n+1,\sigma_e,\sigma}$ and $\sigma_b \le \sigma_n < \sigma_e$ and $c = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$\\
     \hline

### 2019-07-18 popl-2019-journal-v/paper.tex
[POPLjv] Remove macros that leaves out details
[POPLjv] Move encoding function remark (address comment)

@@
 \renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
-
-\renewcommand{\lre}[1][]{\typesetlr{E}^{#1}}
-\renewcommand{\lretrusted}[1][]{\lreg[#1]{\trusted}}
-\renewcommand{\lrexj}[1][]{\typesetlr{E}^{#1}_{\mathrm{xjmp}}}
-\renewcommand{\lrk}[1][]{\typesetlr{K}^{#1}}
-\renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
-\renewcommand{\lrr}[1][\square,\gc]{\lrrg[#1]{}}
-\renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
-\renewcommand{\lro}[1][]{\typesetlr{O}^{#1}}
-\renewcommand{\lrol}{\lro[\preceq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
-\renewcommand{\lror}{\lro[\succeq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
-\renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
-\renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
-\renewcommand{\lrp}{\typesetlr{P}}
-
-\renewcommand{\lrrs}{\typesetlr{R}}
-\renewcommand{\lrm}{\typesetlr{M}}
-\renewcommand{\lrcomp}[1][]{\typesetlr{C}^{#1}}
-\renewcommand{\lrec}[1][]{\typesetlr{EC}^{#1}}
-\renewcommand{\lrheap}[1][]{\typesetlr{H}^{#1}}
 
@@
 It encodes each kind of word differently and all words of the same kind to the same integer.
-In Section~\ref{subsec:overlay-semantics}, we present an overlay semantics which changes the set of words compared to this machine.
-\todo[inline]{Dominique: perhaps delay the discussion of encodetype for the overlay semantics to Section~\ref{subsec:overlay-semantics}?}
-The overlay semantics also has a type encoding function which must encode the new words, so they correspond to their representation on this machine.
-For instance, the encoding of stack-pointer in the overlay semantics should be the same as the encoding of a capability on this machine.
-All the new words in the overlay semantics correspond to a word on this machine, and their encodings should match.
 \end{jversion}
@@
 For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.
+\begin{jversion}
+The new run-time values in the overlay semantics affect the definition of the encoding function $\encType{}$.
+All the new values correspond to concrete capabilities on the \trgcm{} machine which the encoding function must respect.
+For instance, the encoding of a stack pointer in the overlay semantics should be the same as the encoding of a linear capability on the \trgcm{} machine.
+\end{jversion}
 

### 2019-07-18 popl-2019-journal-v/appendix.tex
[POPLjv] Address appendix todos

@@
     \hline
-
   \end{tabular}
@@
    \hline
-   %TODO stk
     \splice{r_1}{r_2}{r_3}                       & \left.
-\sourcecolor\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ &\updReg{r_3}{w_3}\\ &\updReg{r_1}{c})\endarray\right.& \srcalt{$\Phi(r_2) = \stkptr{\perm,\baddr,n,\_}$ and $\Phi(r_3) = \stkptr{\perm,n+1,\eaddr,\aaddr}$ and $\baddr \le n < \eaddr$ and $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$ and $w_2,w_3 = \linCons{\Phi(r_2),\Phi(r_3)}$}\\
+\sourcecolor\updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{0}\\ &\updReg{r_3}{0}\\ &\updReg{r_1}{c})\endarray\right.& \srcalt{$\Phi(r_2) = \stkptr{\perm,\baddr,n,\_}$ and $\Phi(r_3) = \stkptr{\perm,n+1,\eaddr,\aaddr}$ and $\baddr \le n < \eaddr$ and $c = \stkptr{\perm,\baddr,\eaddr,\aaddr}$}\\
     \hline

### 2019-07-18 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Rewrite region future world affine remark

@@
 The third and final rule says that $\spatial$ regions can become $\spatialo$ which indicates that the linear capabilities are affine.
-Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes them affine.
-This is modelled in the world by allowing future worlds to pickup ownership.
+Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes the linear capabilities affine.
+The world a capability is safe with respect to is no longer needed when the capability is dropped, so other worlds may pick up the ownership which the future region relation allows.
 % describe why "Spatial -> spatialo" ~ affine or at least why we have it?
-\todo[inline]{Lau: explain why $\spatialo \future \spatial$ gives us affine capabilitites.
-This is claimed in the technical report, but I don't see how this is different from revoking $\spatial$ regions and adding a new region with the future world relation...
-I have written something, but it does not feel like the story really holds.}
 \begin{figure}[htb]

### 2019-07-18 popl-2019-journal-v/paper.tex
[POPL-jv] Add well-formedness introduction (address comment)

@@
 \subsection{Components, linking, programs, and contexts }
-\label{subsed:components-linking}
+\label{subsec:components-linking}
 The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
@@
 % - Address must be in data memory
-\todo[inline]{Dominique: a bit of intro seems needed here: explain that we require certain syntactic properties corresponding to a linker doing its job properly and this is captured by a a syntactic well-formedness judgement that we explain in this section.}
-
 \begin{jversion}
+  The components introduced in Section~\ref{subsec:components-linking} are pretty much unconstrained.
+  For instance, a component can have multiple linear capabilities for the same piece of memory, and there are no restrictions on seals.
+  In a real system, the operating system and linker  % ???
+  would make sure everything is setup correctly.
+  For instance, they would not allocate multiple linear capabilities for the same memory, and they would ensure sane seal allocation.
+
+  The component notion is used for both trusted and adversarial components.
+  We expect the linker to link the two programs together making sure that all the system invariants are respected.
+  We expect trusted components to be the product of a known compiler that uses \stktokens{} which means that the code must have a certain shape.
+  The proper linker behaviour and the syntactic expectations we have of trusted component are captured in a syntactic well-formedness judgement which we explain in this section.
+
   Informally, a well-formed component respects Write-XOR-Execute.

### 2019-07-18 popl-2019-journal-v/paper.tex
[POPLjv] Remove todo
[POPLjv] Add Stelios reference todo
[POPLjv] Rewrite well-formed component description (address suggestion)

@@
 
-% see e-mail discussion: we do not see a good way to implement this TODO.
-% \todo[inline]{Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.}
+% Update: see e-mail discussion: we do not see a good way to implement this.
+% Dominique: I think it would be less confusing and more in line with the overlay semantics story to not add an extra instruction in Figure~\ref{fig:source-syntax} but rather present it as just a special case of the overlay semantics for the same underlying instructions.
 
@@
 
-  Informally, a well-formed component respects Write-XOR-Execute.
-  This means that there are no execute capabilities for the data memory and no write capability for the code memory.
-  The data memory of a well-formed component respects linearity and it may come with preset data structures.
-  It may, however, not contain capabilities for the code memory and sets of seals.
-  All the available seals of a component should reside in the code memory.
+  The well-formedness judgement imposes a quite rigid structure on well-formed components: a component's code memory may contain only data and sets of seals.
+  The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities (to the component's data memory, sealed with closure seals).
+  This makes a clear separation between code and data memory as neither can contain capabilities for the other.
+  The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking a return seals that were indirectly accessible through the data structure.
+  Informally, a well-formed component respects Write-XOR-Execute which means that all capabilities for code memory at most has read-execute permission and all capabilities for data memory at most has read-write permission.
+  A component's exports should be sealed code or data capabilities, sealed with a closure seal.
 
@@
   This means that each return seal is only used for one call.
-
-  \todo[inline]{Dominique: The following line appears to explain an arbitrary point about well-formedness: I would explain here that well-formedness imposes a quite rigid structure on well-formed components: a component's code memory may contain only data and sets of seals, while the data memory contains only data, non-sealed memory capabilities to the component's data or code memory (respecting linearity) or sealed memory or code capabilities (to the respective part of the component's memory, sealed with closure seals).
-A component's exports should be sealed code or data capabilities, sealed with a closure seal.
-    Also: perhaps move this before the point about adversarial versus trusted components.
-  }
-  Finally, well-formed components (adversarial and trusted) can export what corresponds to closures, i.e.,  capabilities for data and code memory sealed with a closure seal.
 \end{jversion}
@@
 
+% TODO Relate to Stelios' CSF 2019 paper
+
 % \begin{acks}                            %% acks environment is optional

### 2019-07-18 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Fix macros
[POPLjv] Remove todos
I believe all of these have been addressed.

I don't think the intro suggested in the todo is an improvement over the current wording.
[POPLjv] Add intro to permission based condition subsection

@@
-%\todo[inline]{Write small introduction.
-Maybe something about the difficulty of full-abstraction proofs (with references?) but mention that this is easier than some other cases (reference approximate back translation?) because back translation is trivial.}
 % \begin{itemize}
@@
 %solving the recursive domain equation means that we can present the actual definitions
+% Also see the notes from Aina
 In the previous sections, we have sketched what we would like our worlds to look like.
@@
 \label{par:perm-cond}
-% TODO add brief intro to paragraph
+The safe capabilities will be defined by the value relation.
+However, the safety requirements for a capability depends on the authority the capability gives.
+Therefore, rather than bundling everything into the value relation, we first present a number of \emph{permission based conditions} that each spell out what the requirements are for each permission.
+
 The world can be seen as an authority specification which means that it dictates what kind of capabilities can address a certain part of memory.
@@
 
-% TODO introduce read condition
 We capture the essence of what it means for a capability with read permission to be safe in the condition $\readCond{}$.
@@
 
-% TODO add approx square to relation, add gc
 \begin{figure}
@@
 We use this to not relate register content that will be overwritten anyway.
-We write $\lrrname(W)$ to mean $\lrrname(\emptyset)(W)$.
+We write $\lrrs(W)$ to mean $\lrrs(\emptyset)(W)$.
 That is, if we do not need to exclude additional registers, then we simply omit that argument.
@@
   \begin{align*}
-% TODO add approx square
   \lre(W) &= \left\{ \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} \middle| 
@@
       \quad \forall W_\lrrs , W_\lrm \ldotp \\
-      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \wedge\\
+      \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs) \wedge\\
       \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\

### 2019-07-19 popl-2019-journal-v/paper.tex
[POPLjv] Fix macros
A comment.

--- 2019-07-18 popl-2019-journal-v/paper.tex
@@
 \renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
+
+\renewcommand{\lre}[1][]{\typesetlr{E}^{#1}}
+\renewcommand{\lretrusted}[1][]{\lreg[#1]{\trusted}}
+\renewcommand{\lrexj}[1][]{\typesetlr{E}^{#1}_{\mathrm{xjmp}}}
+\renewcommand{\lrk}[1][]{\typesetlr{K}^{#1}}
+\renewcommand{\lrrg}[2][]{\typesetlr{R}^{#1}_{#2}}
+\renewcommand{\lrr}[1][\square,\gc]{\lrrg[#1]{}}
+\renewcommand{\lrrtrusted}[1][]{\lrrg[#1]{\trusted}}
+\renewcommand{\lro}[1][]{\typesetlr{O}^{#1}}
+\renewcommand{\lrol}{\lro[\preceq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
+\renewcommand{\lror}{\lro[\succeq,(\ta,\stkb,\gsigrets,\gsigcloss)]}
+\renewcommand{\lrvg}[2][]{\typesetlr{V}^{#1}_{#2}}
+\renewcommand{\lrvtrusted}[1][]{\lrvg[#1]{\trusted}}
+\renewcommand{\lrp}{\typesetlr{P}}
+
+\renewcommand{\lrrs}{\typesetlr{R}}
+\renewcommand{\lrm}{\typesetlr{M}}
+\renewcommand{\lrcomp}[1][]{\typesetlr{C}^{#1}}
+\renewcommand{\lrec}[1][]{\typesetlr{EC}^{#1}}
+\renewcommand{\lrheap}[1][]{\typesetlr{H}^{#1}}
 
@@
 The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
-More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri_2015}.
+More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri:_2015}.
 These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
@@
 \begin{jversion}
-  This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens_2019}.
+  This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens:_2019}.
   Compared to the earlier text, this version adds X and Y and Z.
@@
 In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
-\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri_2015} with linear capabilities as the main addition.
+\trgcm{} is mainly inspired by CHERI~\citep{watson_cheri:_2015} with linear capabilities as the main addition.
 For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
@@
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri_2015,watson_fast_2016}.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri:_2015,watson_fast_2016}.
 % seals
@@
 It encodes each kind of word differently and all words of the same kind to the same integer.
+In Section~\ref{subsec:overlay-semantics}, we present an overlay semantics which changes the set of words compared to this machine.
+\todo[inline]{Dominique: perhaps delay the discussion of encodetype for the overlay semantics to Section~\ref{subsec:overlay-semantics}?}
+The overlay semantics also has a type encoding function which must encode the new words, so they correspond to their representation on this machine.
+For instance, the encoding of stack-pointer in the overlay semantics should be the same as the encoding of a capability on this machine.
+All the new words in the overlay semantics correspond to a word on this machine, and their encodings should match.
 \end{jversion}
@@
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.
+Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report} for details.
 \begin{figure}[tb]
@@
 For example, stack capabilities in the overlay semantics correspond directly to linear capabilities in the underlying semantics, and they have authority over the part of memory that the overlay views as the stack.
-\begin{jversion}
-The new run-time values in the overlay semantics affect the definition of the encoding function $\encType{}$.
-All the new values correspond to concrete capabilities on the \trgcm{} machine which the encoding function must respect.
-For instance, the encoding of a stack pointer in the overlay semantics should be the same as the encoding of a linear capability on the \trgcm{} machine.
-\end{jversion}
 
@@
 
-Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri_2015} for an overview of previous work.
-The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri_2015,woodruff_cheri_2014}.
+Capability machines originate with \citet{dennis_programming_1966} and we refer to \citet{levy_capability-based_1984} and \citet{watson_cheri:_2015} for an overview of previous work.
+The capability machine formalized in Section~\ref{sec:cap-mach-w-seal-and-lin} is modelled after CHERI~\citep{watson_cheri:_2015,woodruff_cheri_2014}.
 This is a recent, relatively mature capability machine which combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
@@
 
-CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri_2015}.
+CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri:_2015}.
 The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.

### 2019-07-19 popl-2019-journal-v/proving-full-abstraction.tex
A comment.

--- 2019-07-17 popl-2019-journal-v/proving-full-abstraction.tex
@@
 In this presentation, we try to include most of the details.
+\todo[inline]{Dominique: I would try to formulate this more positively.
+It sounds too much like ``In this version, we copy-pasted boring stuff from the TR to try and fill more pages.''.
+We should say \emph{why we think those details are interesting} to the reader.
+Perhaps something like ``In the previous version, we were forced to omit many aspects of the logical relation, and especially the way it deals with the linking model.
+This was unfortunate and therefore the current version takes a bit longer to explain those aspects.
+Readers who are looking for a shorter summary with less detail may ...''}
 We refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
@@
 It also splits the world in three to distribute the authority of the world.
-The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrhname$, $\lrstk[]$, and $\lrfree[]$, respectively. 
-The $\lrhname$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
-The $\lrhname$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
+The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrheap$, $\lrstk[]$, and $\lrfree[]$, respectively. 
+The $\lrheap$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
+The $\lrheap$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
 The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
@@
 \label{def:heap-rel}
-For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrhname$ as:
+For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrheap$ as:
 \[
@@
 Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
-The $\lrhname$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.
-
-To a large extend, the definition of $\lrhname$ is pretty standard.
-$\lrhname$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
+The $\lrheap$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.
+
+To a large extend, the definition of $\lrheap$ is pretty standard.
+$\lrheap$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
 The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction partitions the authority of the world using $\oplus$.
@@
 The heap sub-world contains all seal invariants.
-Similar to memory segments, only one seal invariant should impose restrictions on a seal, which $\lrhname$ makes sure is the case.
+Similar to memory segments, only one seal invariant should impose restrictions on a seal, which $\lrheap$ makes sure is the case.
 \begin{definition}[Free stack relation]
@@
 \end{definition}
-The free stack relation $\lrfree[]$ is in most regards like the heap relation, $\lrhname{}$.
+The free stack relation $\lrfree[]$ is in most regards like the heap relation, $\lrheap{}$.
 The free stack relation, partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under part of the world.
@@
 A value relation based on what it is safe for an adversary to have should prohibit return seals, so such a relation cannot be used to reason about trusted programs.
-For this reason, we define two value relations a trusted $\lrvname_{\trusted}$ and an untrusted $\lrvname$.
+For this reason, we define two value relations a trusted $\lrvtrusted$ and an untrusted $\lrv$.
 Anything safe for unstrusted programs is also safe to give to a trusted program, so the trusted value relation is defined as a super set of the untrusted value relation.
@@
 The untrusted and trusted value relation both relates \srccm{} and \trgcm{} words.
-The untrusted value relation $\lrvname$ relates words that are safe to give to untrusted programs and $\lrvname_\trusted$ relates words that are safe to give to trusted programs.}
+The untrusted value relation $\lrv$ relates words that are safe to give to untrusted programs and $\lrvtrusted$ relates words that are safe to give to trusted programs.}
 \label{fig:value-relation}
@@
 The $\execCond{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
-We later define the $\lrename$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
+We later define the $\lre$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
 An executable capability can have its range of authority shrunk or its current address changed which changes what instructions are executed and thus potentially whether the code respects memory invariants.
@@
 \label{par:untrusted-val-rel}
-The untrusted value relation $\lrvname$ relates all the words that untrusted components can safely posses.
+The untrusted value relation $\lrv$ relates all the words that untrusted components can safely posses.
 That is words that cannot be used to break any memory invariants.
@@
 Further, if we take any sealed capabilities related under $\sigma$ in the seal invariant, then they should produce a safe result.
-We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrename_\mathrm{xjmp}$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
+We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrexj$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
 % Linear and non-linear case. Affects the future world we consider.
@@
 \label{par:trusted-val-rel}
-The trusted value relation $\lrvname_\trusted$ relates everything safe for a trusted component to have without breaking memory invariants.
+% TODO write section
+The trusted value relation $\lrvtrusted$ relates everything safe for a trusted component to have without breaking memory invariants.
 For the most part, we allow them to contain the same words as the untrusted components, but we also need to allow them to have seal sets with trusted closure seals and return seals which we cannot allow untrusted components to have.
@@
 Note that even though the capability has read permission, we do not require the read condition to hold.
-The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrvname$, so the read condition would not hold.
+The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrv[]$, so the read condition would not hold.
 However, trusted code can have access to such seals because we expect the trusted code to treat the seals reasonably.
@@
 
-\caption{The trusted value relation $\lrvname_\trusted$ relates all the words that are safe for trusted components to contain.
+\caption{The trusted value relation $\lrvtrusted$ relates all the words that are safe for trusted components to contain.
 A trusted component may contain untrusted words (\figurename~\ref{fig:untrusted-val-rel}), return seals and trusted closure seals, and code pointers for trusted code.
@@
 We use this to not relate register content that will be overwritten anyway.
-We write $\lrrname(W)$ to mean $\lrrname(\emptyset)(W)$.
+We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$.
 That is, if we do not need to exclude additional registers, then we simply omit that argument.
@@
 \subsubsection{Expression relations}
-The expression relation $\lrename$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e.
+% TODO write section
+The expression relation $\lre$ defines when two capabilities can be used in the $\pcreg$-register to produce related executions, i.e.
 the capabilities can be used to construct configurations in the observation relation.
-The $\lrename$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control flow during execution when a $\tjmp{}$ instruction is executed.
+The $\lre$ relation can be used to reason about the safety of an executable capability, i.e.\ capabilities that can change the control flow during execution when a $\tjmp{}$ instruction is executed.
 In the setting of \srccm{} and \trgcm{}, we can also use sealed capabilities to change the control flow by using the $\txjmp{}{}$ instruction.
-The $\txjmp{}{}$ instruction updates the $\pcreg$ register and the $\rdata$ register, however, the $\lrename$ relation only updates the $\pc$-register, so we cannot use $\lrename$ to reason about sealed capabilities.
-Instead, we define the relation $\lrename_\mathrm{xjmp}$ which relates two pairs of capabilities when they are safe to use with the $\txjmp{}{}$ instruction.
+The $\txjmp{}{}$ instruction updates the $\pcreg$ register and the $\rdata$ register, however, the $\lre$ relation only updates the $\pc$-register, so we cannot use $\lre$ to reason about sealed capabilities.
+Instead, we define the relation $\lrexj$ which relates two pairs of capabilities when they are safe to use with the $\txjmp{}{}$ instruction.
 
@@
 Instead, we need to pick configurations made out of related components, i.e.\ related register files and related memories that respect linearity.
-The type of execution captured by $\lrename$ corresponds to a normal jump.
+The type of execution captured by $\lre$ corresponds to a normal jump.
 When a $\tjmp r$ instruction is interpreted, the $\pcreg$-register is replaced with the contents of register $r$, i.e. the current configuration is plugged with a new pc.
-The $\lrename$ relation is defined in \figurename~\ref{fig:expr-rels}
-
-The $\lrename_\mathrm{xjmp}$ relation looks very much like the $\lrename$ relation.
+The $\lre$ relation is defined in \figurename~\ref{fig:expr-rels}
+
+The $\lrexj$ relation looks very much like the $\lre$ relation.
 It takes related memories and register files (ignoring the $\rdata$ register) and combines them into two configurations.
-Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lroname$ relation.
-The $\lrename_\mathrm{xjmp}$ relation is defined in \figurename~\ref{fig:expr-rels}.
+Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lro$ relation.
+The $\lrexj$ relation is defined in \figurename~\ref{fig:expr-rels}.
 \begin{figure}
@@
 This means that we need to lift relatedness one step further than the components, namely to the level of execution configurations.
-To this end, we define $\lrecname$.
+To this end, we define $\lrec$.
 \begin{definition}[Related execution configuration]
@@
 Definition~\ref{def:exec-conf-rel} essentially says that two executable configurations are related when they are made out of related components.
-That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$ and $\rdata$ registers are related by the $\lrename_\mathrm{xjmp}$ relation.
+That is, the authority of the world must be distributed such that the code and data pointer pairs are safe for execution, i.e.\ the contents of the $\pcreg$ and $\rdata$ registers are related by the $\lrexj$ relation.
 Further, the two memories and the two register files should be related.

### 2019-07-19 popl-2019-journal-v/paper.tex
More comments.

--- 2019-07-18 popl-2019-journal-v/paper.tex
@@
 \renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}
+
+\renewcommand{\lrr}{\lrrg{ }}
 
@@
 \begin{jversion}
-Formally, the judgement $\wdjud[\ta]{\comp}$ specifies what components are well formed, i.e.\ satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
+These requirements are formally expressed by the judgement $\wdjud[\ta]{\comp}$ specifies, i.e.\ it defines when components satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
 The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
@@
 The \textsc{Main} rule requires the component to have an entry point in terms of a main pair from the exports.
-Further, the component should be a base component that satisfies judgement.
+Further, the base component should satisfy the well-formedness judgement.
 The \textsc{Base} judgement ensures sure that the component has a bit of structure.
@@
 \item \textit{Code and data memory are disjoint.}
-\item \textit{Code memory is padded with zeros ($\mspad$).} This prevents the memory of two code memories from being spliced together.
-\item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.} The component judgement ensure that the linear address space of a component is at most referenced by one capability.
-This means that any capability for the linear address space is unique.
+\item \textit{Code memory is padded with zeros ($\mspad$).} This prevents the memory of two code memories from being spliced together.\todo[inline]{Dominique: more explanation needed?}
+\item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.}
+  The component judgement ensure that the linear address space of a component is at most referenced by one capability. This means that any capability for the linear address space is unique.
 \item \textit{There is a partition $A_\var{own}$ of $A_\var{linear}$ that assigns a set of linear-addressed to each address of data memory.} The partitions are used by the component word judgement ($\vdash_\mathrm{comp-word}$) which ensures linear capabilities only address the addresses in the linear address space that is owned by the address where the capability resides in memory.
+  \todo[inline]{Maybe it would be clearer to merge this point and the previous point and explain that all memory can either be addressed by any number of non-linear caps or by at most one linear capability in the component's data memory.}
 \item \textit{All import addresses are part of the data memory.} 

### 2019-07-22 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Implement Dominique's comment about the intro

--- 2019-07-18 popl-2019-journal-v/proving-full-abstraction.tex
@@
 In the next section, we give an overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
+
 \begin{jversion}
-In this presentation, we try to include most of the details.
-We refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
+  In the previous version, we were forced to omit many aspects of the logical relation, and especially the way it deals with the linking model.
+  This was unfortunate as substantial effort goes into this work, and there are insights that may be useful for similar work.
+  Therefore the current version takes a bit longer to explain those aspects.
+  Readers who are looking for a shorter summary with less detail may refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
 \end{jversion}

### 2019-07-22 popl-2019-journal-v/paper.tex
[POPLjv] Added more explanation to code memory padding

@@
 \item \textit{Code and data memory are disjoint.}
-\item \textit{Code memory is padded with zeros ($\mspad$).} This prevents the memory of two code memories from being spliced together.\todo[inline]{Dominique: more explanation needed?}
+\item \textit{Code memory is padded with zeros ($\mspad$).}
+  This prevents code memories from being spliced together.
+  If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow.
+  Further, the possible control-flows of a component suddenly depends on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
 \item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.}

### 2019-07-22 popl-2019-journal-v/paper.tex
[POPLjv] Merge linearity items in comp well-formedness overview

@@
   Further, the possible control-flows of a component suddenly depends on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
-\item \textit{Data address space can be partitioned in a linear part $A_\var{linear}$ and a non-linear part $A_\var{non-linear}$.}
-  The component judgement ensure that the linear address space of a component is at most referenced by one capability. This means that any capability for the linear address space is unique.
-\item \textit{There is a partition $A_\var{own}$ of $A_\var{linear}$ that assigns a set of linear-addressed to each address of data memory.} The partitions are used by the component word judgement ($\vdash_\mathrm{comp-word}$) which ensures linear capabilities only address the addresses in the linear address space that is owned by the address where the capability resides in memory.
-  \todo[inline]{Maybe it would be clearer to merge this point and the previous point and explain that all memory can either be addressed by any number of non-linear caps or by at most one linear capability in the component's data memory.}
+\item \textit{All memory can either be addressed by any number of non-linear capabilities or at most one linear capability.}
+  The data address space is split into $A_\var{non-linear}$ and $A_\var{linear}$ that can be addressed by non-linear capabilities and linear capabilities, respectively.
+  The judgement ensure uniqueness of linear capabilities by letting each address of the data memory take sole ownership of addresses in $A_\var{linear}$.
 \item \textit{All import addresses are part of the data memory.} 

### 2019-07-23 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Revised section about solving the recursive domain equation
First draft.

@@
 \[
-  \Worldh = \RegName \parfun (\World \fun \Regionh + \Regions)
+  \Worldh = \RegName \parfun (\Regionh + \Regions)
 \]
@@
 \[
-  \Worldfs = \RegName \parfun (\World \fun \Regions)
+  \Worldfs = \RegName \parfun \Regions
 \]
@@
 \[
-\Worlds = \RegName \parfun (\World \fun \Regions) \times \Addr 
+\Worlds = \RegName \parfun (\Regions \times \Addr)
 \]
@@
   \begin{array}{l}
-    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
-    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
     \{\revoked \}
@@
   \begin{array}{l}
-    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \uplus \\
-    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \uplus \\
+    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) + \\
+    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) + \\
     \{\revoked \}
@@
 % Also see the notes from Aina
-In the previous sections, we have sketched what we would like our worlds to look like.
+In the previous sections, we sketched what our worlds should be.
 However, the worlds we want constitute a self-referential domain equation for which no solution exists in set theory.
@@
 We move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
-Intuitively, a c.o.f.e. is a set with some extra structure.
+Intuitively, a c.o.f.e.\ is a set with some extra structure.
 Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
@@
 In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
-We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for this particular setting.
+We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for our particular world.
 \begin{definition}[Ordered family of equivalences (o.f.e.)]
-An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and for each $n$ $\nequal$ is an equivalence relation.
-The pair must satisfy the following properties
+An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and $\nequal$ is an equivalence relation for each $n$.
+The pair must satisfy the following properties:
 \begin{enumerate}
@@
 \end{definition}
-\noindent One can think of the indexes on the family of equivalences as a measure of precision.
-The larger $n$ is, the more refined the equivalences are.
-On the other hand as $n$ decreases, the equivalences become more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
+\noindent The index on the family of equivalences is intuitively a measure of precision.
+The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.
+On the other hand as the index decreases, the equivalences become increasingly imprecise and unable to distinguish certain elements.
 At index 0, all precision is lost and the equivalence cannot distinguish anything.
@@
 % TODO write something here about why the different things (the meassure of the ofe and the guaranteed limits of the cofe) gives us in terms of finding a solution.
-Functions between c.o.f.e.'s must retain the added structure relatively to sets.
-To this end, we require all our functions to be \emph{non-expansive} which means that $n$-equivalences are preserved.
-A function is called \emph{contractive} when it not only retain equivalences but makes them more precise.
+In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
+Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
+This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
+Functions may not only retain equivalences but make it more precise.
+Such a function is said to be \emph{contractive}.
 \begin{definition}
@@
 \noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
-% TODO add comment about standard constructions (x, +, fun, later, etc) and fixed-points (what are the conditions?
-Contractive, and?).
-Mention that it does not suffice to define our worlds, so we need to have some specialized worlds.
-However, this is not the only structure we would like to impose on our sets.
-
-% Not sure this is necessary:
-% % preordered cofe constructions
-% %% cofe to preordered cofe
-% \begin{lemma}[c.o.f.e.'s are preordered c.o.f.e's]
-%   A c.o.f.e $\cofe{X}$, is also a preordered c.o.f.e. with the order $=$.
-% \end{lemma}
-In order to express the domain equation, we must lift all parts of the world definition from sets to c.o.f.e.'s.
-That is, we need to lift the sets of partial functions, monotone non-expansive functions, and relations, respectively to c.o.f.e.'s.
-
-We need to lift the set of relations over $R$, $\Rel{R}$, to a c.o.f.e.
-However, the set of relations have no inherent step-index which means that we need to add one.
-We need to add the step-index in such a way that we can define a family of equivalences that satisfies the c.o.f.e.\ conditions.
-To this end, we define the uniform relations over $R$ $\URel{R}$.
-%% URel
-\begin{definition}[Uniform relation]
-  Given a relation $R \subseteq X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
-  \[
-    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats\ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp (n,r) \in A\}
-  \]
-  The c.o.f.e.\ $\cofe{\URel{R}}$. Define erasure as
-  \[
-    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
-  \]
-  and the family of equivalences as
-  \[
-    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
-  \]
-\end{definition}
-The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step and up and requires the rest to be equal.
-This means that everything is equal at step $0$ as everything is left out.
-
-In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
-In order to have a future world relation that can be used on recursive instances of the world, we need to couple the future world relation to the sets we work with.
-The future world relations is a preorder structure, so we define a notion of \emph{preordered c.o.f.e.'s} which is simply a c.o.f.e.\ with a preorder that preserves the c.o.f.e.\ properties.
-% preordered cofe
+There are a number of standard c.o.f.e.\ constructions for countably infinite sets, products, sums, and functions.
+A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
+The other constructions are used to create c.o.f.e.'s from other c.o.f.e.'s.
+For instance the product construction is defined as follows:
+\begin{definition}[Product c.o.f.e.]
+\label{def:product-cofe}
+\emph{delete me}
+\todo[inline]{insert definition}
+\end{definition}
+The remaining standard constructions are defined similarly.
+Another important construction is the \emph{later} ($\blater$) c.o.f.e.\
+The later c.o.f.e.\ construction is based on one c.o.f.e.\ which it decreases the index of.
+\begin{lemma}[$\blater$ c.o.f.e.]
+  \label{def:later-cofe}
+  \todo[inline]{replace with actual definition for c.o.f.e (not preoreded!)}.
+  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
+  \[
+    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
+  \]
+  where
+  \[
+    x \nequal[0]_{\blater} x' \wedge x \nequal[n]_{\blater} x' \text{ iff } x \nequal[n+1]_X x'
+  \]
+\end{lemma}
+The later c.o.f.e.\ is used to make sure that the index decreases before the self-reference which gives something to recurse on in order to find a solution for the domain equation.
+If we only use the standard constructions to construct a self-referential equation and make sure to guard every negative occurrence of the self reference with the later c.o.f.e.\ construction, then the equation has a solution.
+\todo[inline]{Insert the solution theorem here?}
+
+Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
+The worlds we sketched have additional structure in terms of a future world relation.
+The future world relation is captured by adding a preorder to c.o.f.e.'s:
 \begin{definition}[Preordered c.o.f.e.]
@@
 \end{definition}
-With the preordere c.o.f.e.\ definition, we can define the remainder constructs necessary to define the world we need.
-Any c.o.f.e.\ can trivially be lifted to a preordered c.o.f.e.\ where the ordering is the identity.
-
-The region invariants use monotone, non-expansive functions, so we need a c.o.f.e.\ to represent them.
-This c.o.f.e.\ is defined as follows.
-%% Monotone, non-expansive
-\begin{lemma}[The c.o.f.e. of monotone and non-expansive functions]
+The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
+There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
+We will need the standard construction for later, product, and sum later, so we define them here.
+The reader may skip these definitions for now and refer back to them later.
+\begin{definition}[$\blater$ preordered c.o.f.e.]
+  \label{def:later-p-cofe}
+  \todo[inline]{replace with actual definition for preordered c.o.f.e}.
+  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
+  \[
+    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
+  \]
+  where
+  \[
+    x \nequal[0]_{\blater} x' \wedge x \nequal[n]_{\blater} x' \text{ iff } x \nequal[n+1]_X x'
+  \]
+\end{definition}
+We will also need the canonical preordered c.o.f.e.\ for sum.
+\begin{definition}[Product preordered c.o.f.e.]
+  \label{def:product-p-cofe}
+  \todo[inline]{Add definition}
+
+\end{definition}
+\begin{definition}[Sum preordered c.o.f.e.]
+  \label{def:sum-p-cofe}
+  \todo[inline]{Add definition}
+
+\end{definition}
+
+We cannot rely entirely on the standard constructions to create the worlds we need.
+Instead, we need to define c.o.f.e.'s and preordered c.o.f.e.'s that meet our needs.
+
+Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
+
+In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get the world equation that we want.
+First, we define the c.o.f.e.\ that we need for the memory invariants.
+Note that we do not need a preorder on this.
+As a reminder, we sketched the memory invariants to as follows
+\[
+  \World \monfun \Rel{\MemSeg \times \MemSeg}
+\]
+We will need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
+The world will be given to us as the solution to the recursive equation.
+Generally, we can lift a set of relations to a c.o.f.e.\ by adding an index to it that is downwards closed.
+This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
+%% URel
+\begin{definition}[Uniform relation]
+  \label{uniform-rel-cofe}
+  Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
+  \[
+    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
+  \]
+  The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
+  \[
+    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
+  \]
+  where erasure is defined as
+  \[
+    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
+  \]
+\end{definition}
+The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step and up and requires the rest to be equal.
+This means that everything is equal at step $0$ as everything is left out.
+
+Next, we define a c.o.f.e. of monotone, non-expansive functions
+\begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
+\label{def:monnefun-cofe}
   Given a c.o.f.e.
-$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future\right)$ where $\future$ is a preorder on $W$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ , define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ where
+$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
   \[
@@
   \]
-\end{lemma}
-Next, we define the product preordered c.o.f.e.\
-%% Product
-\begin{lemma}[Product preordered c.o.f.e.]
-  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ define the product preordered c.o.f.e.\ as $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
-  where the family of equivalences are defined as follows
-  \[
-    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
-  \]
-  and the preorder $\future$ is defined as follows
-  \[
-    (x',y') \future (x,y) \text{ iff } x' \future x \wedge y' \future y
-  \]
-\end{lemma}
-The product c.o.f.e.\ is used to combine the sub-worlds into the complete world.
-The regions are also defined as a product, however, the preorder associated with the regions should correspond to the future region relation we sketched in \sectionname~\ref{subsubsec:ft-and-revocation}.
-Shared regions use the canonical ordering, but the canonical ordering does not suffice for spatial regions.
-Therefore, we define a preordered c.o.f.e.\ specifically for the construction of the spatial region tags.
-\begin{lemma}[Spatial region preordered c.o.f.e.]
-  Given a preordered c.o.f.e.\ $\left( H , \left( \nequal_H \right)_{n=0}^\infty , \future_H\right)$ the following is a preordered c.o.f.e.\ \[\left( \{\spatial,\spatial\} \times H + \{\pure\}, \nequal, \future \right)\]
-  where the family of equivalences $\left( \nequal \right)_{n=0}^\infty$ is the canonical family of equivalences and $\future$ is defined as the future region relation from \sectionname~\ref{subsubsec:ft-and-revocation}.
-\end{lemma}
-
-Similarly to the product preordered c.o.f.e. we have a sum preordered c.o.f.e.\ which is used to construct the heap sub-world.
-%% Sum
-\begin{lemma}[Sum preordered c.o.f.e.]
-  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ define the sum preordered c.o.f.e.\ as $\left(X \cup Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
-  where the family of equivalences are defined as follows
-  \[
-    a \nequal b \text{ iff }
-    \left\{
-      \begin{aligned}
-        a,b \in X \wedge a \nequal_X b \vee{} \\
-        a,b \in Y \wedge a \nequal_Y b
-      \end{aligned}
-    \right.
-  \]
-  and the preorder
-  \[
-    a \future b \text{ iff }
-    \left\{
-      \begin{aligned}
-        a,b \in X \wedge a \future_X b \vee{} \\
-        a,b \in Y \wedge a \future_Y b
-      \end{aligned}
-    \right.
-  \]
-\end{lemma}
-\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
-%% Partial function from set
-\begin{lemma}[The preordered c.o.f.e.\ of partial functions]
-  Given a set $S$ and a preordered c.o.f.e $\left( P, \left( \nequal_P \right)_{n=0}^{\infty}, \future_P \right)$, the preordered c.o.f.e.\ of partial non-expansive functions $\left(S \parfun P, \left( \nequal  \right)_{n=0}^\infty,\future\right)$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
-  \[
-     f \nequal g \text{ iff } 
-     \left\{
-       \arraycolsep=0pt \array{l}
-         n = 0 \vee{} \\
-         \dom(f) = \dom(g) \wedge \forall x \in \dom(f) \ldotp f(x) \nequal_P g(x)
+\end{definition}
+In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\, and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
+
+Next, we construct the seal invariant.
+As a reminder, we sketched the seal invariant to be
+\[
+  \Seal \parfun \World \monnefun \URel{\SealableCaps\times\SealableCaps}
+\]
+In order to define this as a c.o.f.e., we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
+This means that we just need to lift the set of seals to a c.o.f.e.\ and define the c.o.f.e.\ partial functions.
+Generally, all countable sets can be lifted to a c.o.f.e.\ with a family of equivalences that has full precision for all indices except index 0.
+\begin{definition}[Sets to c.o.f.e.]
+  \label{def:lift-set-cofe}
+  A countable set $S$ can be lifted to a c.o.f.e. $\cofe{S}$ where
+  \begin{itemize}
+  \item $\nequal[0]$ is the total relation
+  \item for $n > 0$ the relation $\nequal$ is the normal equality on the set
+  \end{itemize}
+\end{definition}
+This definition can be used to lift the set of all seals $\Seal$ to a c.o.f.e.
+
+The c.o.f.e. of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions
+\begin{definition}[The c.o.f.e. of partial, non-expansive functions]
+\label{def:parfun-cofe}
+  Given a c.o.f.e.
+$\left(S,\left(\nequal[n]_S\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(H,\left(\nequal[n]_H\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(S \parfun H,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
+  \[
+    h \nequal h' \text{ iff } \dom(h) = \dom(h') \wedge \forall w \in \dom(h) \ldotp h(w) \nequal_H h'(w)
+  \]
+\end{definition}
+The seal invariant is defined using Definition~\ref{def:lift-set-cofe} to lift $\SealableCaps$ to a c.of.e.\, defining a c.o.f.e.\ similar to the one for memory invariants, and finally combining the two with Definition~\ref{def:parfun-cofe}.
+
+Next, we want to define the regions of the world.
+As mentioned, regions have the added structure of a future region relation, so we need to define them as a preordered c.o.f.e.
+The future region relation was presented in Section~\ref{subsubsec:ft-and-revocation}, so we just need to use it for our definitions.
+The future world relation for pure region is trivial as it is just equality
+The pure regions we sketched are the easiest to define in terms of a c.o.f.e.\ as the future relation is just equality.
+\begin{definition}[Pure region preordered c.o.f.e.\ ]
+  \label{def:pure-reg-p-cofe}
+  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of pure regions as
+  \[
+  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
+  \]
+  where the preorder $\future$ is equality, and the family of equivalences distributes to the underlying c.o.f.e.'s, i.e.
+given $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
+  \[
+  (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
+  \]
+\end{definition}
+Definition~\ref{def:pure-reg-p-cofe} gives us the pure region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
+
+Next, we define the preordered c.o.f.e.\ for shared regions.
+The future region relation on shared regions is non-trivial, so the definition of the preordered c.o.f.e.\ is a bit more involved.
+\begin{definition}[Shared region preordered c.o.f.e.\ ]
+  \label{def:shared-reg-p-cofe}
+  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
+  \[
+    \Regions = \{\spatial\}\times \var{MemInv} + \{\spatialo\} \times \var{MemInv'} + \{\revoked\}
+  \]
+  define the preordered c.o.f.e.\ of shared regions as
+  \[
+  \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
+  \]
+  where the preorder $\future$ is defined as
+  \begin{mathpar}
+  \inferrule{ r \in \Regions }{ r \future r }
+  \and
+  \inferrule{ }{ \revoked \future (\spatial,\_)}
+  \and
+  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
+  \end{mathpar}
+  and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
+  for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
+  \begin{itemize}
+  \item $n=0$
+  \item $n > 0$ and one of the following is true
+    \begin{itemize}
+    \item $r = r' = \revoked$
+    \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
+    \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
+    \end{itemize}
+  \end{itemize}
+\end{definition}
+Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:shared-reg-p-cofe} define the preoreded c.o.f.e.'s necessary for the regions in our worlds.
+
+Next, we need to define the three sub-worlds for heap, encapsulated stack frames, and free stack.
+The definitions of the sub-worlds are very similar, so we can capture the essence of all three sub worlds in one definition.
+\begin{definition}[Sub-world preordered c.o.f.e.]
+  \label{def:sub-world-p-cofe}
+  Given preordered c.o.f.e. $\left(\var{Reg},\left(\nequal[n]_\var{Reg}\right)_{n=0}^{\infty},\future_\var{Reg}\right)$ define the sub-world preordered c.o.f.e.\ as
+  \[
+    \left(\nats \parfun \var{Reg},\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \]
+  where the family of equivalences is defined by given $W,W' \in \nats \parfun \var{Reg}$,
+  \[
+    W \nequal W' \text{ iff } \dom(W) = \dom(W') \wedge \forall r \in \dom(W)\ldotp W(r) \nequal W'(r)
+  \]
+  and the future world relation is defined by given $W,W' \in \nats \parfun \var{Reg}$,
+  \[
+  W' \future W \text{ iff } \left\{
+    \array{l}
+    \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
+    \quad \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'.i(m(r)) \future W(r)
      \endarray
@@
 \]
-  and the canonical preorder
-  \[
-    f \future g \text{ iff } \dom(f) = \dom(g) \wedge \forall x \in \dom(f) \ldotp f(X) \future_P g(x)
-  \]
-\end{lemma}
-
-Functions between preordered c.o.f.e's must preserve the ordering which means that they must be monotone with respect to the ordering.
-% monotone functions
-\begin{definition}
-  \label{def:monotone-preordered-cofe}
-  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty},\future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} ,\future_Y\right)$ be two preordered c.o.f.e.'s and $f$ a function from the set $X$ to the set $Y$.
-  The function $f$ is \emph{monotone} when for all $x, x' \in X$,
-\[
-  x' \future_X x \implies f(x) \future_Y f(x')
-\]
-\end{definition}
-
-% TODO a preordered c.o.f.e. w/ our world?
-
-In some sense, the step-index in a c.o.f.e. gives us a metric that we may use to find a fixed-point.
-However, for a fixed-point to exist, the metric must decrease before a self-reference.
-As this is not an inherent property of c.o.f.e.'s, we have a c.o.f.e.\ construction called \emph{later} with the sole purpose of decrementing the index.
-%% Later cofe
-\begin{lemma}[$\blater$ preordered c.o.f.e.]
-  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
-  \[
-    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
-  \]
-  where
-  \[
-    x \nequal[0]_{\blater} x' \wedge x \nequal[n]_{\blater} x' \text{ iff } x \nequal[n+1]_X x'
-  \]
-\end{lemma}
-The later needs to guard all negative self-references in a recursive domain equation in order for a solution to exist.
-The modality does not have to be placed directly on the self-reference.
-For instance in the Theorem~\ref{thm:rec-dom-eq-sol}, it is placed on the world itself.
-The later placement does, however, affect what the logical relation definitions in Section~\ref{subsec:logical-relation} looks like, so in the end it is a matter of taste.
-% Solution theorem
+\end{definition}
+The three sub-worlds we need can be obtained from Definition~\ref{def:sub-world-p-cofe}.
+First, we define the free stack sub-world simply by using the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) with the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
+The result is the sub-world for the free stack
+\[
+  \Worldfs = \nats \parfun \Regions
+\]
+
+Next, we define the call stack sub-world.
+To this end, we use Definition~\ref{def:shared-reg-p-cofe} to get the shared region.
+We need to combine this with an address.
+We achieve this by first lifting the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combining the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with Definition~\ref{def:product-p-cofe}.
+All in all, we get the a preordered c.o.f.e.\
+\[
+  \Regions \times \Addr
+\]
+which we use with Definition~\ref{def:sub-world-p-cofe} to get the preordered c.o.f.e.\ for the stack sub-world:
+\[
+  \Worlds = \nats \parfun \Regions \times \Addr
+\]
+
+Finally, we define the sub-world for the heap.
+To this end, we combine the pure region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
+The result is a preordered c.o.f.e.\
+\[
+  \Regions + \Regionh
+\]
+which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
+\[
+  \Worldh = \nats \parfun (\Regions + \Regionh)
+\]
+
+With the three sub-worlds, we can combine them to get our worlds using the preordered c.o.f.e.\ construction for products (Definition~\ref{def:product-p-cofe}) which gives us the preordered c.o.f.e.
+\[
+  \Worldh \times \Worlds \times \Worldfs
+\]
+This is the world, we would like to work with, but we still don't have a solution for the recursive domain equation.
+In fact, when we defined the regions, we just specified where the solution to the recursive domain equation should go.
+So what we previously sketched to be the recursive occurrence of the world is now a variable in the above.
+In order to find a solution, we still need to guard the recursive occurrence.
+To this end, we simply use the construction for later preordered c.o.f.e.'s in Definition~\ref{def:later-p-cofe} with the above to get.
+\[
+  \blater (\Worldh \times \Worlds \times \Worldfs)
+\]
+Now the recursive occurrence is guarded which means that there is a solution to the recursive domain equation.
+
 \begin{theorem}
@@
 and
+\todo[inline]{look over relations and make it uniform whether it is $\MemSeg\times\MemSeg$ or $\MemSeg^2$}
 \[
@@
 %% Aleš note? Tutorial note?
-
-The domains in Theorem~\ref{thm:rec-dom-eq-sol} are made from the c.o.f.e.\ and preorder c.o.f.e.\ constructions we presented earlier in this section.
-They are also the definitions we use in our world definition.
-We do not define $\Wor$ to be our worlds.
-Instead, we define our world as
+The solution to the recursive domain equation is the preordered c.o.f.e.\ $\Wor$ in Theorem~\ref{thm:rec-dom-eq-sol}.
+Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
+This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
+This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
 \[
@@
 \]
-which is easier as we only have to use $\xi$ and move under a later when we have gone into the region and need to work with $\Wor$.
+
+%% Lau: not sure whether the following remarks should be included:
+% Note that the later placement could have been different as it just needs to guard the recursive occurrence at some level.
+% It could, for instance, have been directly on the recursive occurrence.
+% By moving the later, we would also move the place we have to go under the later in our later definitions affecting how it is working with them.
+
+% In worlds that has a simple future world relation and no future region relation, it may be an advantage to define the regions in terms of a self-referential equation, solve that, and define the worlds with the solution.
+% The advantage of this approach is that it can be done entirely in c.o.f.e.'s.
 

### 2019-07-23 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Insert missing definitions in cofe section

@@
 \label{def:product-cofe}
-\emph{delete me}
-\todo[inline]{insert definition}
+Given two c.o.f.e.'s $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal \right)_{n=0}^{\infty} \right)$ where the equivalence family is defined as for $(x,y),(x',y') \in X \times Y$
+\[
+  (x,y) \nequal (x',y') \text{ iff } x \nequal x' \wedge y \nequal y'
+\]
 \end{definition}
@@
   \label{def:later-cofe}
-  \todo[inline]{replace with actual definition for c.o.f.e (not preoreded!)}.
-  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
-  \[
-    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
+  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
+  \[
+    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
   \]
@@
   \[
-    x \nequal[0]_{\blater} x' \wedge x \nequal[n]_{\blater} x' \text{ iff } x \nequal[n+1]_X x'
+    x \nequal_{\blater} x' \text{ iff } \left\{
+      \begin{array}{l}
+        n = 0 \vee{}\\
+        x \nequal[n+1]_X x'
+      \end{array}
+    \right.
   \]
@@
   \label{def:later-p-cofe}
-  \todo[inline]{replace with actual definition for preordered c.o.f.e}.
   Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
@@
   \[
-    x \nequal[0]_{\blater} x' \wedge x \nequal[n]_{\blater} x' \text{ iff } x \nequal[n+1]_X x'
+    x \nequal_{\blater} x' \text{ iff } \left\{
+      \begin{array}{l}
+        n = 0 \vee{} \\
+        x \nequal
+        [n+1]_X x'
+      \end{array}
+    \right.
   \]
@@
   \label{def:product-p-cofe}
-  \todo[inline]{Add definition}
-
+  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
+  \[
+  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \]
+  where the preorder $\future$ distributes to the underlying preorder, i.e.\
+  \[
+\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
+  \]
+  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
+  \[
+\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
+  \]
 \end{definition}
@@
   \label{def:sum-p-cofe}
-  \todo[inline]{Add definition}
-
+  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
+  \[
+  \left(X + Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \]
+  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
+  \[
+    \text{for } z,z' \in X + Y,\quad  z' \future z \text{ iff } \left\{
+      \begin{array}{l}
+        z,z' \in X \wedge z' \future_x z \vee{} \\
+        z,z' \in Y \wedge z' \future_Y z
+      \end{array}
+    \right.
+  \]
+  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
+  \[
+    \text{for } z,z' \in X + Y,\quad  z \nequal z' \text{ iff } \left\{
+      \begin{array}{l}
+        z,z' \in X \wedge z \nequal_x z' \vee{} \\
+        z,z' \in Y \wedge z \nequal_Y z'
+      \end{array}
+    \right.
+  \]
 \end{definition}
@@
 \begin{definition}[Uniform relation]
-  \label{uniform-rel-cofe}
+  \label{def:uniform-rel-cofe}
   Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
@@
   \Regionh = 
-  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
-  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps^2})
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
 \end{multline*}
 and
-\todo[inline]{look over relations and make it uniform whether it is $\MemSeg\times\MemSeg$ or $\MemSeg^2$}
 \[
@@
   \begin{array}{l}
-    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg^2}) +\\
+    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) +\\
     \{\revoked\}

### 2019-07-26 popl-2019-journal-v/proving-full-abstraction.tex
Some more comments.

--- 2019-07-22 popl-2019-journal-v/proving-full-abstraction.tex
@@
 In the next section, we give an overview of the relation we define, and then we sketch the full-abstraction proof in Section~\ref{subsec:proof-sketch}.
-
 \begin{jversion}
-  In the previous version, we were forced to omit many aspects of the logical relation, and especially the way it deals with the linking model.
-  This was unfortunate as substantial effort goes into this work, and there are insights that may be useful for similar work.
-  Therefore the current version takes a bit longer to explain those aspects.
-  Readers who are looking for a shorter summary with less detail may refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
+In this presentation, we try to include most of the details.
+\todo[inline]{Dominique: I would try to formulate this more positively.
+It sounds too much like ``In this version, we copy-pasted boring stuff from the TR to try and fill more pages.''.
+We should say \emph{why we think those details are interesting} to the reader.
+Perhaps something like ``In the previous version, we were forced to omit many aspects of the logical relation, and especially the way it deals with the linking model.
+This was unfortunate and therefore the current version takes a bit longer to explain those aspects.
+Readers who are looking for a shorter summary with less detail may ...''}
+We refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
 \end{jversion}
@@
 Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds.
-A world is a semantic model of the memory.
+To a first approximation, a world is a semantic model of the memory.
 In its simplest form, it is a collection of invariants that the memory must satisfy.
@@
 That is, for every address in the two memory segments, the words that reside there must be related.
-In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some the intuition.
+In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some intuition.
 If we have two integers, then they are related when they are equal.
@@
 The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
-In order to model this uniqueness, we need to keep track of which parts of memory that are uniquely referenced and make sure that only one linear capability references the unique parts.
+In order to model this uniqueness, we need to keep track of which parts of memory are uniquely referenced and make sure that only one linear capability references the unique parts.
 %% Bookkeeping done by worlds
@@
 %%% spatial/spatial_owned
+\todo[inline]{Dominique: Maybe it would be clearer to rename spatial\_owned/spatial to spatial/spatial\_shadow or something similar.
+Spatial\_owned is really the default meaning of spatial, and what we've called spatial is a peculiar thing that we use for a very specific purpose.
+  This would make the discussion easier to follow, I think.}
+
 We cannot let multiple linear capabilities reference the same memory, so we add ownership to spatial regions.
@@
 \]
+\todo[inline]{Dominique: Perhaps rename pure to shared?  Or introduce the term pure earlier on?}
 For readability, we also add a tag $\pure$ to the shared regions:
@@
 \]
+\todo[inline]{Dominique: $+$?  Why not $\cup$?}
 \srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
@@
 This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions.
+\todo[inline]{Dominique: step missing in the reasoning: expressing that the capability is not linear is done by stripping the world ...
+\emph{and then} requiring that the capability is valid w.r.t.\ this stripped world.}
 We refer to this as the pure part of the world and define the function $\purePart{}$ which removes all ownership.
@@
 \label{subsubsec:seals}
+\todo[inline]{
+  I would try to link back to the discussion of worlds as collections of assumptions by first explaining that correctness of values may depend on another type of assumptions than those about the contents of memory, namely the intended use case of seals.
+  Therefore, in addition to a relation on memory segments, some regions also carry a seal interpretation function that etc.
+  I think this would make the role of the seal interpretation function clearer.
+  }
 % StkTokens designate seals for certain purposes, return, closure, etc
@@
 The designation of return seal is a system invariant that is modelled in the world as a \textit{seal interpretation function}.
-This function takes a seal and returns a world-indexed relation that relates all the sealables that may be sealed with this specific seals.
+This function takes a seal and returns a world-indexed relation that relates all the sealables that may be sealed with this specific seal.
 That is
@@
 The seal interpretation function should be an invariant, so we need to add it to a type of region that is invariant.
+\todo[inline]{It's not clear at this point why it would be problematic to add it to a non-invariant region.
+  I would add a discussion saying something like ``In StkTokens, once a seal has been used for a specific purpose (e.g.
+for sealing return capability pairs for a specific call site), it can never be reused for a different purpose.
+This is because there may still be copies of return capabilities out there, signed with the seal.
+  This situation is similar to the situation for non-linear memory capabilities, so we only allow shared regions to carry seal interpretation functions, as we will see that those regions can never be revoked in future worlds.''
+}
 Shared regions are invariant, so we add the seal interpretation function to it:
@@
 % Memory evolves over time, we need to model this.
-During machine execution, the memory contents changes and parts of memory may be repurposed.
+During machine execution, the memory contents changes and parts of memory may be repurposed\todo[inline]{Dominique: Add something like ``, at least when we can be sure that there are no capabilities left in the system whose validity depends on the old contents of the memory.''?}.
 The world must model such memory changes.
@@
 To this end, we have a future world relation $\future$ that describes how worlds and thus memory may change over time.
+\todo[inline]{Rather than introducing the future world relation as ``how memory may change over time'', then its precise definition and then monotonicity, I would explain monotonicity first and the precise definition of the future world relation second.
+  IMHO, it makes a lot more sense if you think about the future world as modelling precisely those changes in assumptions that do not invalidate the relatedness of values.
+  In other words, I would say something like ``Very often, it is the case that relatedness of two capabilities doesn't change if extra assumptions in the system are added.
+For example, two related capabilities remain related when an extra invariant is added on unrelated memory, or when a stack frame that it does not reference is dropped.
+  In Kripke logical relations, this kind of assumption changes, that do not invalidate the relatedness of values, are modelled by the future world relation.
+  etc.''
+  }
+
+  \todo[inline]{Maybe explain here that Kripke future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
+    However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on the capability anymore.
+    By linking revocation of regions immediately to linearity, it will require less explanation later that only spatial regions can be revoked.
+  }
 In order to relate \trgcm{} and \srccm{}, we are mainly interested in changes to the stack as it repurposes stack frames.
-Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions that govern call frames that have been popped from the call stack, so the stack frame can be reused for a new call frame.
+Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions for call frames that are popped from the call stack, to make place for a new frame.
 We allow introduction of new regions simply by extension.
@@
 Things are a bit more complicated when it comes to revocation.
-\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).}
-For technical reasons, we need the worlds to be extensional, so we cannot simply forget about regions that we want to remove.
+\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).
+}
+\dominique{To be honest, I don't actually know what exactly would go wrong if we simply removed revoked regions.
+  Perhaps we should think this through?
+}
+For technical reasons, we need the worlds to be extensional\todo{Dominique: meaning?}, so we cannot simply forget about regions that we want to remove.
 We keep removed regions in the world and tag them as removed with a $\revoked$ tag.
 This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.
-Not all regions that should be revocable.
+
+Not all regions should be revocable.
+\todo[inline]{
+  Perhaps explain intuition with a phrase like ``Since regions express assumptions that the validity of other values in the system may depend on, we should only be able to revoke a region when we are sure that nothing depends on it.
+  This is never the case for shared regions, as blabla copying blabla...
+  However, if we have a linear capability to a block of memory, then linearity ensures that no other capability in the system can reference the same memory.''
+  }
+
 Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities, in some sense, can be revoked.
@@
 
+\todo[inline]{Dominique: it would be great if we could add an example here of two concrete capabilities, one of which is linear, the worlds in which they're valid and explain how we can revoke the linear one and repurpose the memory it addresses, but still keep the other capability valid in the new world?
+}
+
 With the future region relation in place, we define the future world relation as follows: For worlds $W$ and $W'$,
@@
 \]
-The relation says that each of the three worlds must be an extension of the past world and each of the existing region must have a future region.
+The relation says that each of the three worlds must be an extension of the past world and each of the existing regions must have a future region.
 % injective function
 Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
-The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.
+The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.\dominique{I wonder if the renaming function obviates the need for the ``revoked'' trick.}
 
@@
 The memory invariants describe what certain parts of memory looks like.
-In future worlds, part of the memory may change but for the memory invariant to be invariant, it should keep relating the same memory segments even when the rest of the world changes.
+In future worlds, part of the memory may change but for the memory invariant to be invariant, it should continue relating the same memory segments even when the rest of the world changes.
+\dominique{This discussion might be easier if we had talked about monotonicity earlier.}
 For this reason, we require the world-indexed memory invariants to be monotone in the world.
@@
 \label{subsubsec:rec-dom-eq}
+\dominique{Explain that this section provides technical details on how the recursive world equation can be solved and that readers may skip this if they like.}
 % TODO consider new structure (the below)
@@
 Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
+\dominique{mention that this is a standard technique (cite the relevant papers) and an abstracted form of step-indexing (cite some more papers).}
 We move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
@@
 \end{definition}
+\dominique{Explain how an o.f.e. can be seen as a set with extra structure.}
 \noindent One can think of the indexes on the family of equivalences as a measure of precision.
+\dominique{Introduce phrases like ``up to n steps'' etc.}
 The larger $n$ is, the more refined the equivalences are.
@@
   Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
-  We 
   The function $f$ is
@@
 \end{definition}
-With the preordere c.o.f.e.\ definition, we can define the remainder constructs necessary to define the world we need.
+With the preordered c.o.f.e.\ definition, we can define the remainder constructs necessary to define the world we need.
 Any c.o.f.e.\ can trivially be lifted to a preordered c.o.f.e.\ where the ordering is the identity.
@@
   \]
+  \dominique{shouldn't this be $x \nequal[n+1]_{\blater} x' \text{ iff } x \nequal[n]_X x'$}
 \end{lemma}
@@
 For instance in the Theorem~\ref{thm:rec-dom-eq-sol}, it is placed on the world itself.
-The later placement does, however, affect what the logical relation definitions in Section~\ref{subsec:logical-relation} looks like, so in the end it is a matter of taste.
+The later placement does, however, affect what the logical relation definitions in Section~\ref{subsec:logical-relation} looks like, so in the end it is a matter of taste.\dominique{This reasoning is odd: it's a matter of taste because it affects what the LR looks like?}
 % Solution theorem
@@
 \end{definition}
-The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible (e.g., the regions are not overlapping), but when we use a world, we will always make sure that it is sensible by using memory satisfaction.
+The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible (e.g., the regions are not overlapping).
+Only when we know that a certain memory satisfies the world (as a memory specification, see Section~\ref{subsubsec:mem-sat}), will we be sure that the world's specifications are actually non-contradictory.
 
@@
 The memory satisfaction relation makes sure that this is the case, so in a sense the memory satisfaction also acts as a well-formedness judgement for worlds.
+
+\dominique{At this point, the paper is starting to read like a tech report.
+  Are we sure it's a good idea to provide this much technical detail?
+  I would propose to not include all definitions, but a representative selection and refer to the TR for full detail.
+}
 
@@
 The logical relation we define ends up as a cyclic definition.
-The circularity is resolved by another use of step-indexing in the definitions, but the circularity also poses a chicken and egg kinda situation with respect to the order in which the definitions of the relations should be presented.
+The circularity is resolved by another use of step-indexing in the definitions, but the circularity also poses a chicken and egg problem with respect to the order in which the definitions of the relations should be presented.
 There is no canonical way of presenting the logical relation as we are bound to make forward references.
-For this reason, we suggest making a cursorily first read through to get an overview followed by a more thorough read.
+For this reason, we suggest making a cursory first read through to get an overview followed by a more thorough read.
 % TODO something about relation naming?

### 2019-07-26 popl-2019-journal-v/proving-full-abstraction.tex

--- 2019-07-23 popl-2019-journal-v/proving-full-abstraction.tex
@@
 \end{definition}
+\dominique{Explain how an o.f.e. can be seen as a set with extra structure.}
 \noindent The index on the family of equivalences is intuitively a measure of precision.
+\dominique{Introduce phrases like ``up to n steps'' etc.}
 The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.

### 2019-07-27 popl-2019-journal-v/paper.tex
Some more comments.
[POPLjv] Update terminology for regions to spatial, shadow, and shared

@@
 
+\renewcommand{\spatialo}{\mathrm{spatial}}
+\renewcommand{\spatial}{\mathrm{shadow}}
+\renewcommand{\pure}{\mathrm{shared}}
+
+\renewcommand{\purePart}[1]{\plainfun{sharedPart}{#1}}
+
 \newcommand{\xjmpres}[1]{\plainfun{xjmpRes}{#1}}
@@
 
+\renewcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
+\renewcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
+\renewcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+
 \section{Introduction}
@@
 Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be built.
+\dominique{Reference Patrignani, Ahmed and Clarke's survey paper?}
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
-
-% TODO Relate full-abstraction to other notions of secure compilation
-
-% TODO Relate to Stelios' CSF 2019 paper
+Our results do not apply to dynamic compromise scenarios because they are
+intended to be used in the verification of a secure compiler where these
+scenarios are not relevant.
+\dominique{Perhaps worth pointing out that (I suspect that) our step-bounded reasonability and FTLR, actually does apply in dynamic compromise scenario's.
+The idea would be to use the theorem multiple times with different step bounds and a different boundary between trusted and untrusted code.
+  More concretely, if a component A starts behaving non-reasonable after 10 steps, then first apply the FTLR with n = 9 and component A trusted.
+  Then apply the FTLR again with n > 10 and component A no longer trusted.
+  Then you get stronger well-bracketedness guarantees until n=9 and weaker ones after that.
+  This is a step-bounded version of the idea explained in \cite{patrignani_modular_2016}.}
+
+\dominique{Relate full-abstraction to other notions of secure compilation}
+
+\dominique{Relate to Stelios' CSF 2019 paper}
+
+\dominique{Relate to Thomas Van Strydonck's ICFP19 paper which uses linear capabilities (at a higher level) to enforce separation logic in a fully abstract way.}
 

### 2019-07-27 popl-2019-journal-v/proving-full-abstraction.tex
Readd a comment that got lost in the merge...
[POPLjv] Correct later cofe and preordered cofe definition
[POPLjv] Update terminology for regions to spatial, shadow, and shared

@@
 On the other hand, if a memory segment is governed by a spatial region, then only a linear capability may reference it.
+
 %%% spatial/spatial_owned
-\todo[inline]{Dominique: Maybe it would be clearer to rename spatial\_owned/spatial to spatial/spatial\_shadow or something similar.
-Spatial\_owned is really the default meaning of spatial, and what we've called spatial is a peculiar thing that we use for a very specific purpose.
-  This would make the discussion easier to follow, I think.}
-
-We cannot let multiple linear capabilities reference the same memory, so we add ownership to spatial regions.
-Only the spatial owned regions can be referenced by a capability.
-Even though only spatial owned regions can be referenced by linear capabilities, we still have unclaimed spatial regions that specify what part of memory looks like.
-When there is an unclaimed spatial region, then there may be a compatible world  that claims the ownership of that spatial region which means that the memory may still be referenced (we return to the notion of compatible worlds w.r.t.
+We cannot let multiple linear capabilities reference the same memory.
+The spatial region represent ownership a part of memory which ensures that a linear capability is not aliased.
+Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
+To this end, we have shadow regions.
+A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
+This does not mean that the memory is necessarily not referenced.
+A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t.
 ownership in Section~\ref{subsubsec:joining-worlds}).
-Specifically, we add tags $\spatial$ and $\spatialo$ to the spatial regions:
+Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
 \[
@@
   \begin{array}{l}
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
     \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
-    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
   \end{array} \right.
 \]
-\todo[inline]{Dominique: Perhaps rename pure to shared?  Or introduce the term pure earlier on?}
 For readability, we also add a tag $\pure$ to the shared regions:
@@
 \emph{and then} requiring that the capability is valid w.r.t.\ this stripped world.}
-We refer to this as the pure part of the world and define the function $\purePart{}$ which removes all ownership.
-\begin{definition}[The pure part of a world]
+We refer to this as the shared part of the world and define the function $\purePart{}$ which turns all spatial regions into shadow copies.
+\begin{definition}[The shared part of a world]
   \label{def:purePart}
@@
     However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on the capability anymore.
-    By linking revocation of regions immediately to linearity, it will require less explanation later that only spatial regions can be revoked.
+    By linking revocation of regions immediately to linearity, it will require less explanation later that only shadow regions can be revoked.
   }
@@
   \[
-    x \nequal_{\blater} x' \text{ iff } \left\{
+    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
       \begin{array}{l}
         n = 0 \vee{}\\
-        x \nequal[n+1]_X x'
+        x \nequal_X x'
       \end{array}
@@
   \[
-    x \nequal_{\blater} x' \text{ iff } \left\{
+    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
       \begin{array}{l}
@@
         x \nequal
-        [n+1]_X x'
+        [n]_X x'
       \end{array}
@@
 The future region relation was presented in Section~\ref{subsubsec:ft-and-revocation}, so we just need to use it for our definitions.
-The future world relation for pure region is trivial as it is just equality
-The pure regions we sketched are the easiest to define in terms of a c.o.f.e.\ as the future relation is just equality.
-\begin{definition}[Pure region preordered c.o.f.e.\ ]
+The future world relation for the shared region is trivial as it is just equality
+The shared regions we sketched are the easiest to define in terms of a c.o.f.e.\ as the future relation is just equality.
+\begin{definition}[Shared region preordered c.o.f.e.\ ]
   \label{def:pure-reg-p-cofe}
-  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of pure regions as
+  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of shared regions as
   \[
@@
 \end{definition}
-Definition~\ref{def:pure-reg-p-cofe} gives us the pure region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
+Definition~\ref{def:pure-reg-p-cofe} gives us the shared region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
 
@@
 Finally, we define the sub-world for the heap.
-To this end, we combine the pure region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
+To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
 The result is a preordered c.o.f.e.\
@@
 We do not know when the executable capability will be used, so it should be safe even in the future when the memory has changed.
-The $\purePart{}$ function strips the world of all spatial ownership.
+The $\purePart{}$ function turns the spatial regions of a world into shadow copies.
 This means that the capability cannot depend on linear capabilities and thus the contents of the stack.
@@
 However, this only goes for non-linear words.
-For linear capabilities, it should only be in future worlds of the pure part of $W$.
+For linear capabilities, it should only be in future worlds of the shared part of $W$.
 % TODO write why we have the two cases. (it allows us to have the "other" capability by the one we consider - and it feels like the only reason)

### 2019-07-27 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Change sum to cup

@@
   \begin{array}{l}
-    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
-    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\ 
+    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})  \\
   \end{array} \right.
@@
 \[
-  \Worldh = \RegName \parfun (\Regionh + \Regions)
-\]
-\todo[inline]{Dominique: $+$?  Why not $\cup$?}
+  \Worldh = \RegName \parfun (\Regionh \cup \Regions)
+\]
 \srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
@@
   \begin{array}{l}
-    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
-    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
+    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\
+    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\
     \{\revoked \}
@@
   \begin{array}{l}
-    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) + \\
-    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) + \\
+    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
+    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
     \{\revoked \}
@@
 There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
-We will need the standard construction for later, product, and sum later, so we define them here.
+We will need the standard construction for later, product, and sum, so we define them here.
 The reader may skip these definitions for now and refer back to them later.
@@
 \end{definition}
-\begin{definition}[Sum preordered c.o.f.e.]
+\begin{definition}[Union preordered c.o.f.e.]
   \label{def:sum-p-cofe}
@@
   \[
-  \left(X + Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \left(X \cup Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
   \]
@@
   \[
-    \text{for } z,z' \in X + Y,\quad  z' \future z \text{ iff } \left\{
+    \text{for } z,z' \in X \cup Y,\quad  z' \future z \text{ iff } \left\{
       \begin{array}{l}
@@
   \[
-    \text{for } z,z' \in X + Y,\quad  z \nequal z' \text{ iff } \left\{
+    \text{for } z,z' \in X \cup Y,\quad  z \nequal z' \text{ iff } \left\{
       \begin{array}{l}
@@
   \[
-    \Regions = \{\spatial\}\times \var{MemInv} + \{\spatialo\} \times \var{MemInv'} + \{\revoked\}
+    \Regions = \{\spatial\}\times \var{MemInv} \cup \{\spatialo\} \times \var{MemInv'} \cup \{\revoked\}
   \]
@@
 Finally, we define the sub-world for the heap.
-To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
+To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the union preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
 The result is a preordered c.o.f.e.\
 \[
-  \Regions + \Regionh
+  \Regions \cup \Regionh
 \]
@@
 \[
-  \Worldh = \nats \parfun (\Regions + \Regionh)
+  \Worldh = \nats \parfun (\Regions \cup \Regionh)
 \]
@@
 \[
-  \Worldh = \RegionName \parfun (\Regions + \Regionh)
+  \Worldh = \RegionName \parfun (\Regions \cup \Regionh)
 \]
@@
   \begin{array}{l}
-    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) +\\
+    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \cup\\
     \{\revoked\}

### 2019-07-27 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] implement comment about sharedWorld

@@
 Given a world, we want to be able to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
-This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions.
-\todo[inline]{Dominique: step missing in the reasoning: expressing that the capability is not linear is done by stripping the world ...
-\emph{and then} requiring that the capability is valid w.r.t.\ this stripped world.}
+This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions and then requiring that the capability is valid w.r.t.\ this new world.
 We refer to this as the shared part of the world and define the function $\purePart{}$ which turns all spatial regions into shadow copies.

### 2019-07-28 popl-2019-journal-v/discussion.tex
Some more remarks.

@@
   This goes to show how we must consider what high-level language we want in order to answer the question of what properties we must enforce to get a fully-abstract compiler.
+  \dominique{Yes, but what if we want exceptions, can StkTokens be modified to support it?
+What about continuations?
+Also: should we mention the limited additional control flow we seem to get if we omit the stack base check (where  adversaries can use a kind of non-preemptive multi-threading)?
+Perhaps the answer to some of these question is that we don't know but we should still give it.}
   % Lau: Say something about address hiding?
@@
   However, fully-abstract compilers do not extend vertically in the sense that if we add a new feature in the target and source language, then the full-abstraction theorem does not follow immediately.
+  \dominique{I find this ``vertical extensionality'' both vague and unrealistic. Should we really discuss this?}
   Intuitively, this is the case because the new feature may have an unforeseen interaction with existing features that breaks a desired property.
@@
   % new paragraph: conditional full-abstraction
+  \dominique{This paragraph should perhaps also point out the link to the dynamic compromise paper?}
   Our full-abstraction theorem, Theorem~\ref{thm:full-abstraction}, is not pure full-abstraction as it requires the components to be reasonable and well-formed.
   In other words, if we were to define a compiler phase that targets \srccm{}, then we would also have to show that every program it generates is well-formed and reasonable in order to use the full-abstraction result.
+  \dominique{The word ``however'' is wrong in the next sentence.
+IIUC, you're explaining *why* we require the previous compiler phase to have ensured reasonability instead of enforcing reasonability in StkTokens.
+  This is very confusing.}
+  \dominique{Also: I'm not sure it's a good idea to lump together reasonability and well-formedness.
+    In my understanding, well-formedness is a set of *syntactic* constraints and the requirement that the linker has distributed seals properly.
+    As such, the discussion below seems to be only about reasonability, not well-formedness.
+  }
   However, even though the well-formed and reasonable conditions are fairly simple, it is too late to ensure them at this point of the compilation.
@@
 \stktokens{} does not support this as it expects to get a stack pointer for the entire stack it handed out.
+\dominique{Err... I don't see the problem.
+  Passing stack references to callees should work fine, as long as the callee treats them linearly (i.e.\ doesn't try to copy them).
+  }
 However, if we have allocated part of the stack for a different purpose and handed out a reference for it, then the stack pointer cannot reference the same memory as the stack pointer is a linear capability.
+\dominique{I don't understand this sentence.}
 However, while it is not possible to allocate a reference for the stack, one can easily imagine that the stack could still be used to pass and return values.
+\dominique{Or this one...}
 \end{jversion}
+
+\dominique{Perhaps worth mentioning the problem with adding support for linear values in a real C compiler: how to explain to the compiler that it should treat the linear values linearly and not use optimizations that will break this.}
 

### 2019-07-29 popl-2019-journal-v/discussion.tex
Add a discussion of the implementation of continuations and exceptions
Not sure what, if anything, to write about: the last question.

@@
   A compiler is secure when it enforces the properties of high-level languages which begs the question what properties should we enforce.
-  When it comes to fully-abstract compilation, then the answer is that all the properties of the high-level language should be enforced which means that the real question is what high-level language we would like.
+  When it comes to fully-abstract compilation, then the answer is that all the properties of the high-level language should be enforced, so the real question is what high-level language we would like.
   \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/CC, then we need to come up with a different enforcement scheme.
@@
   This goes to show how we must consider what high-level language we want in order to answer the question of what properties we must enforce to get a fully-abstract compiler.
-  \dominique{Yes, but what if we want exceptions, can StkTokens be modified to support it?
-What about continuations?
-Also: should we mention the limited additional control flow we seem to get if we omit the stack base check (where  adversaries can use a kind of non-preemptive multi-threading)?
-Perhaps the answer to some of these question is that we don't know but we should still give it.}
   % Lau: Say something about address hiding?
+
+  We conjecture that \stktokens{} can support a limited form of continuations and exceptions.
+  Specifically, the continuations and exceptions would be passed to the callee as sealed code capabilities.
+  In order for the continuations and exceptions to work with the stack, they would have to be sealed with the return seals.
+  This requirement puts two strong limitations on continuations and exceptions.
+  First of all, they can be used by the immediate callee, but they cannot be used in subsequent calls made by the callee (assuming they use \stktokens{}).
+  The problem is that the callee would have to take up part of the stack for their local stack frame.
+  This would break up the return token that should be used with the continuation (or exception).
+  If the callee really needs to pass on the continuation, then they would have to construct a wrapper that unravels the stack and restores the necessary return token.
+  Second, a call that passes continuations or exceptions in this way has to pass the same continuations and exceptions in every call.
+  The problem is that just like the return code capability, the continuation and exception capabilities are non-linear.
+  This means that a callee can store them for subsequent calls preventing us from using different continuations and exceptions for different calls.
+
+  \dominique{Also: should we mention the limited additional control flow we seem to get if we omit the stack base check (where  adversaries can use a kind of non-preemptive multi-threading)?}
 
@@
   Fully-abstract compilers compose vertically which means that if we defined a compiler phase that targets \srccm{}, then we could use the full-abstraction theorem immediately.
-  However, fully-abstract compilers do not extend vertically in the sense that if we add a new feature in the target and source language, then the full-abstraction theorem does not follow immediately.
+  However, fully-abstract compilers do not extend horizontally in the sense that if we add a new feature in the target and source language, then the full-abstraction theorem does not follow immediately.
   \dominique{I find this ``vertical extensionality'' both vague and unrealistic. Should we really discuss this?}

### 2019-07-29 popl-2019-journal-v/discussion.tex
Rewrite discussion about reasonable/well-formed
Split the discussion in two, so they are not lumped together.

@@
 
+\begin{thesisonly}
   % Say something about horizontal vs vertical composition.
@@
   However, fully-abstract compilers do not extend horizontally in the sense that if we add a new feature in the target and source language, then the full-abstraction theorem does not follow immediately.
-  \dominique{I find this ``vertical extensionality'' both vague and unrealistic. Should we really discuss this?}
   Intuitively, this is the case because the new feature may have an unforeseen interaction with existing features that breaks a desired property.
@@
   After all, the capabilities we use here do not hide the addresses in any meaningful way, so a new ``address hiding'' capability would be necessary to enforce this property.
-  If we extended \trgcm{} and \srccm{} with new such capabilities, then we would have to redo the entire full-abstraction proof.
+  If we extended \trgcm{} and \srccm{} with such new capabilities, then we would have to redo the entire full-abstraction proof.
+\end{thesisonly}
 
@@
   In other words, if we were to define a compiler phase that targets \srccm{}, then we would also have to show that every program it generates is well-formed and reasonable in order to use the full-abstraction result.
-  \dominique{The word ``however'' is wrong in the next sentence.
-IIUC, you're explaining *why* we require the previous compiler phase to have ensured reasonability instead of enforcing reasonability in StkTokens.
-  This is very confusing.}
-  \dominique{Also: I'm not sure it's a good idea to lump together reasonability and well-formedness.
-    In my understanding, well-formedness is a set of *syntactic* constraints and the requirement that the linker has distributed seals properly.
-    As such, the discussion below seems to be only about reasonability, not well-formedness.
-  }
-  However, even though the well-formed and reasonable conditions are fairly simple, it is too late to ensure them at this point of the compilation.
-  For instance, reasonability require that return seals are not passed to the adversary in calls.
-  In order to ensure this with \stktokens{}, we would have to add checks to the calling convention that ensures that seals are not passed as arguments in calls.
-  In other words, we would need to add protection to the trusted program against itself which of course shouldn't be necessary.
-  The calling convention should only include protection schemes that protects against untrusted programs.
-  The compiler should generate programs that are actually trustworthy, so we do not have to add protection against the generated programs.
-  However, when we consider \srccm{} components, we have already committed to a program consisting of low-level instructions.
-  The correct place to ensure that our programs are well-formed and reasonable is when we actually generate the instructions and commit to the low-level code which is the phase that goes from the high-level language (or some intermediate representation) to a low-level machine like \srccm{}.
+  Without the reasonability constraint, \stktokens{} would have to enforce reasonability instead.
+  That is, \stktokens{} would have to dynamically ensure that no return seals or means to get return seals are passed in calls.
+  Essentially, such checks would protect the trusted code against itself which shouldn't be necessary in the first place.
+  Instead, the compiler should generate reasonable code that never exhibits the unreasonable behaviour.
+  This should be done in a compiler phase where more information about the original program is available, e.g.\ the compilation phase that commits to the low-level translation.
+  Similarly for the syntactic constraints given by well-formedness.
+  The compiler should make sure to generate code that satisfies the well-formedness judgement, so it can be executed by the machine.
 

### 2019-07-29 popl-2019-journal-v/proving-full-abstraction.tex
Rewrite seal interpretation section a bit

@@
 \label{subsubsec:seals}
-\todo[inline]{
-  I would try to link back to the discussion of worlds as collections of assumptions by first explaining that correctness of values may depend on another type of assumptions than those about the contents of memory, namely the intended use case of seals.
+\todo[inline]{ I would try to link back to the discussion of worlds as collections of assumptions by first explaining that correctness of values may depend on another type of assumptions than those about the contents of memory, namely the intended use case of seals.
   Therefore, in addition to a relation on memory segments, some regions also carry a seal interpretation function that etc.
   I think this would make the role of the seal interpretation function clearer.
-  }
-% StkTokens designate seals for certain purposes, return, closure, etc
-In order to guarantee well-bracketed control flow and local-state encapsulation, \stktokens{} designates seals for return capabilities and closures and relies on designated seals only being used for their intended purposes.
-The designation of return seal is a system invariant that is modelled in the world as a \textit{seal interpretation function}.
-This function takes a seal and returns a world-indexed relation that relates all the sealables that may be sealed with this specific seal.
-That is
+} So far, the world represents a collection of assumptions on the memory contents that value correctness may depend on.
+However, value correctness may depend on other assumptions.
+Specifically, \stktokens{} has certain assumption on the seals used for return capabilities and closures.
+For instance, return seal must only be used to seal the return pointer of one specific return point.
+Therefore, in addition to a relation on memory segments, some regions also carry a \emph{seal interpretation function} that relates the sealables that may be sealed with a given seal.
   \[
@@
 %% Shared region has a seal indexed invariant that specifies what sealables can be sealed with a specific seal
-In Section~\ref{subsubsec:ft-and-revocation}, we will see that not all regions are invariant.
-The seal interpretation function should be an invariant, so we need to add it to a type of region that is invariant.
-\todo[inline]{It's not clear at this point why it would be problematic to add it to a non-invariant region.
-  I would add a discussion saying something like ``In StkTokens, once a seal has been used for a specific purpose (e.g.
-for sealing return capability pairs for a specific call site), it can never be reused for a different purpose.
+  In \stktokens{}, once a seal has been used for a specific purpose (e.g.\ for sealing return capability pairs for a specific call site), it can never be reused for a different purpose.
   This is because there may still be copies of return capabilities out there, signed with the seal.
-  This situation is similar to the situation for non-linear memory capabilities, so we only allow shared regions to carry seal interpretation functions, as we will see that those regions can never be revoked in future worlds.''
-}
-Shared regions are invariant, so we add the seal interpretation function to it:
+  This situation is similar to the situation for non-linear memory capabilities, so we only allow shared regions to carry seal interpretation functions, as we will see that those regions can never be revoked in future worlds.
 \begin{multline*}

### 2019-07-29 popl-2019-journal-v/proving-full-abstraction.tex
Add comment about skipping the recursive domain equation section

@@
 \label{subsubsec:rec-dom-eq}
-\dominique{Explain that this section provides technical details on how the recursive world equation can be solved and that readers may skip this if they like.}
 % TODO consider new structure (the below)
@@
 Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
-In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.
+In this section we present some of the technical details on solving as well as the final world domain equation (Theorem\ref{thm:rec-dom-eq-sol}).
+This section may be skipped by readers with no interest in the recursive domain equation.
 

### 2019-07-29 popl-2019-journal-v/proving-full-abstraction.tex
Implement comment about ofes as sets with structure

@@
 \end{definition}
-\dominique{Explain how an o.f.e. can be seen as a set with extra structure.}
-\noindent The index on the family of equivalences is intuitively a measure of precision.
-\dominique{Introduce phrases like ``up to n steps'' etc.}
+\noindent An o.f.e.\ can be thought of as a set with extra structure.
+Intuitively, the index on the family of equivalences is a measure of precision.
 The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.
@@
 At index 0, all precision is lost and the equivalence cannot distinguish anything.
+\dominique{Introduce phrases like ``up to n steps'' etc.}
 % TODO: Lau write intuition for sequences and why they are necessary.

### 2019-07-30 popl-2019-journal-v/proving-full-abstraction.tex
Restructure and rewrite future world section

@@
 \end{multline*}
-
-We will also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it is unambiguous.
+We also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it is unambiguous.
 
@@
 % Memory evolves over time, we need to model this.
-During machine execution, the memory contents changes and parts of memory may be repurposed\todo[inline]{Dominique: Add something like ``, at least when we can be sure that there are no capabilities left in the system whose validity depends on the old contents of the memory.''?}.
-The world must model such memory changes.
-% Future World
-To this end, we have a future world relation $\future$ that describes how worlds and thus memory may change over time.
-\todo[inline]{Rather than introducing the future world relation as ``how memory may change over time'', then its precise definition and then monotonicity, I would explain monotonicity first and the precise definition of the future world relation second.
-  IMHO, it makes a lot more sense if you think about the future world as modelling precisely those changes in assumptions that do not invalidate the relatedness of values.
-  In other words, I would say something like ``Very often, it is the case that relatedness of two capabilities doesn't change if extra assumptions in the system are added.
+Very often, relatedness of two capabilities does not change if extra assumptions in the system are added.
 For example, two related capabilities remain related when an extra invariant is added on unrelated memory, or when a stack frame that it does not reference is dropped.
-  In Kripke logical relations, this kind of assumption changes, that do not invalidate the relatedness of values, are modelled by the future world relation.
-  etc.''
-  }
-
-  \todo[inline]{Maybe explain here that Kripke future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
-    However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on the capability anymore.
-    By linking revocation of regions immediately to linearity, it will require less explanation later that only shadow regions can be revoked.
-  }
-In order to relate \trgcm{} and \srccm{}, we are mainly interested in changes to the stack as it repurposes stack frames.
-Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions for call frames that are popped from the call stack, to make place for a new frame.
-We allow introduction of new regions simply by extension.
-That is, a future world can have new regions that was not present in a past world.
-Things are a bit more complicated when it comes to revocation.
-\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).
-}
-\dominique{To be honest, I don't actually know what exactly would go wrong if we simply removed revoked regions.
-  Perhaps we should think this through?
-}
-For technical reasons, we need the worlds to be extensional\todo{Dominique: meaning?}, so we cannot simply forget about regions that we want to remove.
-We keep removed regions in the world and tag them as removed with a $\revoked$ tag.
-This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.
-
-Not all regions should be revocable.
-\todo[inline]{
-  Perhaps explain intuition with a phrase like ``Since regions express assumptions that the validity of other values in the system may depend on, we should only be able to revoke a region when we are sure that nothing depends on it.
-  This is never the case for shared regions, as blabla copying blabla...
-  However, if we have a linear capability to a block of memory, then linearity ensures that no other capability in the system can reference the same memory.''
-  }
-
-Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities, in some sense, can be revoked.
-A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory which makes it safe to repurpose.
-On the other hand, normal capabilities may have an alias, so it is not safe to repurpose the memory a normal capability has authority over.
-Only spatial regions specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
-The $\revoked$ region does not specify any memory, so we drop the world indexed relation.
-That is, we have
+In Kripke logical relations, this kind of assumption changes, that do not invalidate the relatedness of values, are modelled by the future world relation $\future$.
+The future world relation can also be thought of as the model of memory changes over time.
+
+The memory invariants in the world should be defined, so they are monotone with respect to the future world relation.
+This means that the system assumptions capabilities rely on for safety are defined such that capabilities remain safe during execution.
+Specifically, we require the world-indexed memory invariants to be monotone in the world.
+This means that a pair of memory segments that are related now will stay related in future worlds.
+This changes the spatial regions as follows:
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
+    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
+    \{\revoked \}
+  \end{array} \right.
+\]
+We make a similar change to $\Regionh$, and we make the seal invariant monotone as well.
+
+Kripke future world relation must be defined such that future worlds are extensions of the past world.
+In other words, future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
+However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on assumption anymore.
+Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it sure to repurpose the memory.
+In terms of the world, this corresponds to dropping a region and replacing it with a new one.
+Due to the fact that future worlds must be an extension of the past world, we cannot drop a region.
+Instead, we mark it as revoked which for all intends and purposes corresponds to dropping the region.
+
+We add a $\revoked$ tag to the spatial regions $\Regions$ which results in
 \[
@@
 \]
-The rules for the future region relation are displayed in Figure~\ref{fig:ft-reg-rel}.
-% Revocation of spatial regions. 'revocation' of linear capabilities.
-The first rule says that all regions are future regions of themselves, so regions are allowed to stay unchanged.
-The second rule says that $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
-The third and final rule says that $\spatial$ regions can become $\spatialo$ which indicates that the linear capabilities are affine.
+The $\spatialo$-region signifies that a linear capability \emph{may} depend on it which means that it cannot be revoked.
+On the other hand, no capability can depend on a $\spatial$-region, so it can be safely revoked.
+This may be confusing as a $\spatial$-region does not mean that you ``have'' the capability, but we just explained that having a linear capability means that you can repurpose the memory it points to.
+As we will see later, the logical relation makes disjoint splits of ownership, so each capability gets its own ownership to depend on.
+A linear capability that is repurposed will get an entirely new world to depend on.
+After all, a future world would have to respect the old assumptions, but a repurposed capability needs to depend on new assumptions.
+The new world constructed for the repurposed linear capability would have a revoked region in place of the $\spatialo$-region, but the new world would not be a future world of the world it depended on before.
+While the repurposed capability gets a new world to depend on, all other capabilities in the system must remain valid in a future world to the one they depend on.
+Further, the future worlds must be compatible (agree on everything but ownership) with the new world.
+This means that all the $\spatial$-regions must be replaced with $\revoked$ regions which is why $\spatial$ regions can be turned into $\revoked$ regions.
+
+We define the future world relation in terms of a future region relation which is displayed in Figure~\ref{fig:ft-reg-rel}.
+Apart from being revoked, a region can stay the same, or a $\spatial$ region can become $\spatialo$ which models the affinity of the linear capabilities.
 Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes the linear capabilities affine.
@@
 \end{figure}
-
-\todo[inline]{Dominique: it would be great if we could add an example here of two concrete capabilities, one of which is linear, the worlds in which they're valid and explain how we can revoke the linear one and repurpose the memory it addresses, but still keep the other capability valid in the new world?
-}
-
-With the future region relation in place, we define the future world relation as follows: For worlds $W$ and $W'$,
+The above repurposing discussion is exemplified in \figurename~\ref{fig:world-example}.
+The Figure displays two capabilities $c_{\var{normal}}$ and $c_{\var{linear}}$ that are normal and linear, respectively.
+The linear capability is valid with respect to $W_1$ which has the necessary $\spatialo$ region, and the normal capability is valid with respect to $W_2$ which has a $\pure$ region.
+The two worlds are compatible as $W_2$ has a $\spatial$ region that matches the $\spatialo$ region of $W_1$.
+When the linear capability is repurposed, it must be reflected by the worlds $W_1'$ and $W_2'$.
+In both new worlds, the $r_2$ region is replaced with a $\revoked$ region.
+$W_1'$ has a new $\spatialo$ region at $r_3$, and $W_2'$ gets a matching $\spatial$ region.
+The new world for the linear capability, $W_1'$, is not a future world of $W_1$ as it replaces a $\spatialo$ region with a $\revoked$ region which is not allowed by the future region relation.
+The new world for the normal capability, $W_2'$, is a future world of $W_2$ as the future region relation allows $\spatial$ regions to be revoked.
+The normal capability $c_{\var{normal}}$ remains valid in $W_2'$ as the $\pure$ region it depends on is monotone with respect to the future world relation.
+\begin{figure}\centering
+
+  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+    \begin{scope}
+  \scope
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+      \foreach \x in {.5,1,...,10}
+      {
+        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {};
+      };
+  \endscope
+  \draw [decorate,decoration={brace,amplitude=7pt,raise=4pt},yshift=0pt] (-.025,0.5) -- (-.125,5) node (lin) [black,midway,xshift=-0.25cm,align=center] { };
+  \draw (-5,3) edge[-latex, bend right, in=180, out=-20, dashed] node[below] {$c_{\var{lin}}$}(lin);
+
+  \draw [decorate,decoration={brace,amplitude=7pt,raise=4pt},yshift=0pt] (-.025,6) -- (-.125,10.5) node (norm) [black,midway,xshift=-0.25cm,align=center] { };
+  \draw (-5,8.5) edge[-latex, bend right, in=180, out=-20] node[below] {$c_{\var{norm}}$} (norm);
+
+  \draw (-0.25,10.5) edge[very thin, densely dashed, opacity=6] (11,10.5);
+  \draw (-0.25,6) edge[very thin, densely dashed, opacity=6] (11,6);
+  \draw (-0.25,5) edge[very thin, densely dashed, opacity=0.5] (11,5);
+  \draw (-0.25,0.5) edge[very thin, densely dashed, opacity=0.5] (11,0.5);
+\end{scope}
+
+  \begin{scope}[shift={(11,0)}]
+  \scope
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+      \foreach \x in {.5,1,...,10}
+      {
+        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {};
+      };
+  \endscope
+
+  \draw (-4,11.5) node {\huge $W_1$};
+  \draw (-4,8) node (w1r1) {$r_1$};
+  \draw (-4,5.5) node (w1r2) {$r_2$};
+
+  \draw[red, fill=red, fill opacity=.5] (-.5,10.5) rectangle (1.5,6) node[fitting node] (w1s) { };
+  \draw (w1r1) edge[->,in=180, out=10] (w1s.west);
+
+  \draw[blue, fill=blue, fill opacity=.6] (-.5,5) rectangle (1.5,0.5) node[fitting node] (w1l) { };
+  \draw (w1r2) edge[->,in=180, out=-30] (w1l.west);
+
+  \draw (8.5,11.5) node {\huge $W_2$};
+  \draw (8.5,8) node  (w2r1) {$r_1$};
+  \draw (8.5,5.5) node (w2r2) {$r_2$};
+
+  \draw[red, fill=red, fill opacity=.5] (3,10.5) rectangle (5,6) node[fitting node] (w2s) { };
+  \draw (w2r1) edge[->,in=0, out=170] (w2s.east);
+
+  \draw[gray, fill=gray, fill opacity=.4] (3,5) rectangle (5,0.5) node[fitting node] (w2l) { };
+  \draw (w2r2) edge[->,in=0, out=210] (w2l.east);
+
+  \end{scope}
+
+  \begin{scope}[shift={(11,-15)}]
+  \scope
+    \clip (-.1,-.1) rectangle (4.6,11.1);
+    \fill[fill=white] (0,0) rectangle (4.5,11);
+    \draw (0,0) -- (0,11);
+    \draw (4.5,0) -- (4.5,11);
+      \foreach \x in {.5,1,...,10}
+      {
+        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {};
+      };
+  \endscope
+
+  \draw (-4,11.5) node (w1) {\huge $W_1'$};
+  \draw node[above of=w1,rotate=-90] {\huge $\not\sqsubseteq$};
+  \draw (-4,8) node (w1r1) {$r_1$};
+  \draw (-4,5.5) node (w1r2) {$r_2$};
+  \draw (-4,3) node (w1r3) {$r_3$};
+
+  \draw[red, fill=red, fill opacity=.5] (-.5,10.5) rectangle (1.5,6) node[fitting node] (w1s) { };
+  \draw (w1r1) edge[->,in=180, out=10] (w1s.west);
+
+  % TODO do something about the capabilities
+  \node[draw,rectangle, black, minimum size=2mm] (w1r) at (-2,5) {  };
+  \node[fit=(w1r),black,inner sep=-\pgflinewidth,cross out, draw] (test) { };
+  \draw (w1r2) edge[->, out=0] (w1r);
+
+  \draw[green, fill=green, fill opacity=.7] (-.5,5) rectangle (1.5,0.5) node[fitting node] (w1l) { };
+  \draw (w1r3) edge[->,in=180, out=-30] (w1l.west);
+
+  \draw (8.5,11.5) node (w2) {\huge $W_2'$};
+  \draw node[above of=w2,rotate=-90] {\huge $\sqsubseteq$};
+  \draw (8.5,8) node  (w2r1) {$r_1$};
+  \draw (8.5,5.5) node (w2r2) {$r_2$};
+  \draw (8.5,3) node (w2r3) {$r_3$};
+
+  \draw[red, fill=red, fill opacity=.5] (3,10.5) rectangle (5,6) node[fitting node] (w2s) { };
+  \draw (w2r1) edge[->,in=0, out=170] (w2s.east);
+
+  \node[draw,rectangle, black, minimum size=2mm] (w2r) at (6.5,5) {  };
+  \node[fit=(w2r),black,inner sep=-\pgflinewidth,cross out, draw] (test) { };
+  \draw (w2r2) edge[->, out=180,in=45] (w2r);
+
+  \draw[gray, fill=gray, fill opacity=.4] (3,5) rectangle (5,0.5) node[fitting node] (w2l) { };
+  \draw (w2r3) edge[->,in=0, out=210] (w2l.east);
+
+  \end{scope}
+
+   \matrix [draw,above right, column sep=1.5mm,xshift=.4cm,yshift=.4cm] at (current bounding box.south west) {
+  &\node[red, fill=red, fill opacity=.5,minimum size=0.4cm,label=right:{\footnotesize Shared region}] {}; \\
+    &\node[gray, fill=gray, fill opacity=.4,minimum size=0.4cm,label=right:{\footnotesize Shadow region}] {}; \\
+     \node[green, fill=green, fill opacity=.7,minimum size=0.4cm] {};
+    &\node[blue, fill=blue, fill opacity=.6,minimum size=0.4cm,draw,label=right:{\footnotesize Spatial region}] {}; \\
+    &\node[draw,rectangle, black, minimum size=4mm] (rev) {};
+    \node[fit=(rev),black,inner sep=-\pgflinewidth,cross out, draw, label=right:{\footnotesize Revoked region}] {};\\
+    \node[minimum size=0.4cm] (ln) {}; & \node[minimum size=0.4cm,label=right:{\footnotesize Normal capability}] (rn) {};\\
+    \node[minimum size=0.4cm] (ll) {}; & \node[minimum size=0.4cm,label=right:{\footnotesize Linear capability}] (rl) {};\\
+};
+\draw (ln.west) edge[-latex,out=-20,in=160] (rn.east);
+\draw (ll.west) edge[-latex,out=-20,in=160,dashed] (rl.east);
+
+%   \matrix [draw,below right, column sep=1.5mm, anchor=south west] at (current bounding box.south west) {
+%     &\node[red, fill=red, fill opacity=.5,minimum size=0.4cm,label=right:{\footnotesize Shared region}] {}; \\
+%     &\node[gray, fill=gray, fill opacity=.4,minimum size=0.4cm,label=right:{\footnotesize Shadow region}] {}; \\
+%      \node[green, fill=green, fill opacity=.7,minimum size=0.4cm] {};
+%     &\node[blue, fill=blue, fill opacity=.6,minimum size=0.4cm,draw,label=right:{\footnotesize Spatial region}] {}; \\
+%     &\node[draw,rectangle, black, minimum size=2mm] (rev) {};
+%      \node[fit=(rev),black,inner sep=-\pgflinewidth,cross out, draw, label=right:{\footnotesize Revoked region}] {};\\
+% };
+\end{tikzpicture}
+
+\caption{Example what happens to worlds when a linear capability $c_{\var{lin}}$ is repurposed.
+  The linear capability is originally valid with respect to $W_1$.
+  In $W_1'$ there is a new spatial region that the linear capability is valid with respect to (the different coloured regions signify different invariants).
+  $W_1'$ still has $r_2$, but now it points to a $\revoked$ region.
+  This means that $W_1'$ is not a future world of $W_1$.
+  The normal capability $c_{\var{norm}}$ is originally valid with respect to $W_2$ and should stay valid with respect to the new world $W_2'$.
+The $W_2'$ is a future world of $W_2$ as the $\spatial$ region in $W_2$ is replaced with a $\revoked$ region which is permitted by the future world relation.}
+\label{fig:world-example}
+\end{figure}
+
+With the future region relation in place, we can define the future world relation as follows: For worlds $W$ and $W'$,
 \[
@@
 The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.\dominique{I wonder if the renaming function obviates the need for the ``revoked'' trick.}
-
-The memory invariants are world dependent, but the worlds may change according to the future world relation, so what does that mean for the memory invariants?
-The memory invariants describe what certain parts of memory looks like.
-In future worlds, part of the memory may change but for the memory invariant to be invariant, it should continue relating the same memory segments even when the rest of the world changes.
-\dominique{This discussion might be easier if we had talked about monotonicity earlier.}
-For this reason, we require the world-indexed memory invariants to be monotone in the world.
-This means that a pair of memory segments that are related now will stay related in future worlds.
-This changes the spatial regions as follows:
-\[
-  \Regions = \left\{
-  \begin{array}{l}
-    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
-    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
-    \{\revoked \}
-  \end{array} \right.
-\]
-We make a similar change to $\Regionh$.
-% TODO update terminology: memory invariant, H, and seal invariant, H_\sigma
 
@@
 
+\lau{TODO: insert monotonicity lemma}
 \paragraph{The trusted value relation}
@@
 \end{figure}
+
 \subsubsection{Register file relation}

### 2019-07-30 popl-2019-journal-v/paper.tex
Some more remarks.
Some more suggestions for related work.
Rewrite discussion about reasonable/well-formed
Split the discussion in two, so they are not lumped together.
Formatting
Restructure and rewrite future world section
[POPLjv] Add survey paper citation

@@
 \input{preamble-illu}
+\usetikzlibrary{fit}
+\makeatletter
+\tikzset{
+  fitting node/.style={
+    inner sep=0pt,
+    fill=none,
+    draw=none,
+    reset transform,
+    fit={(\pgf@pathminx,\pgf@pathminy) (\pgf@pathmaxx,\pgf@pathmaxy)}
+  },
+  reset transform/.code={\pgftransformreset}
+}
+\makeatother
 
@@
     {\begin{comment}}{\end{comment}}
+\newenvironment{thesisonly}{}{}
 
@@
 \todo[inline]{TODOS ENABLED!!!}
-\todo[inline]{Lau: for camera ready version update reference to technical report (and upload the technical report).}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016}), but a real secure compiler is yet to be built.
-\dominique{Reference Patrignani, Ahmed and Clarke's survey paper?}
+Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016, patrignani_2019}), but a real secure compiler is yet to be built.
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are
-intended to be used in the verification of a secure compiler where these
-scenarios are not relevant.
+Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
 \dominique{Perhaps worth pointing out that (I suspect that) our step-bounded reasonability and FTLR, actually does apply in dynamic compromise scenario's.

### 2019-07-30 popl-2019-journal-v/paper.tex
[POPLjv] Update todos

@@
 \begin{jversion}
+  % TODO add outline of new work
   This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens_2019}.
   Compared to the earlier text, this version adds X and Y and Z.
-  \todo{new version of TR?}
   This paper is accompanied by a technical report~\citep{technical_report_popl} with the elided details and proof.
@@
 
-\dominique{Relate full-abstraction to other notions of secure compilation}
-
-\dominique{Relate to Stelios' CSF 2019 paper}
-
-\dominique{Relate to Thomas Van Strydonck's ICFP19 paper which uses linear capabilities (at a higher level) to enforce separation logic in a fully abstract way.}
+\dominique{TODO Relate full-abstraction to other notions of secure compilation}
+
+\dominique{TODO Relate to Stelios' CSF 2019 paper}
+
+\dominique{TODO Relate to Thomas Van Strydonck's ICFP19 paper which uses linear capabilities (at a higher level) to enforce separation logic in a fully abstract way.}
 

### 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Small changes

@@
 \label{subsubsec:seals}
-\todo[inline]{ I would try to link back to the discussion of worlds as collections of assumptions by first explaining that correctness of values may depend on another type of assumptions than those about the contents of memory, namely the intended use case of seals.
-  Therefore, in addition to a relation on memory segments, some regions also carry a seal interpretation function that etc.
-  I think this would make the role of the seal interpretation function clearer.
-} So far, the world represents a collection of assumptions on the memory contents that value correctness may depend on.
+So far, the world represents a collection of assumptions on the memory contents that value correctness may depend on.
 However, value correctness may depend on other assumptions.
@@
 
-  % TODO do something about the capabilities
   \node[draw,rectangle, black, minimum size=2mm] (w1r) at (-2,5) {  };
@@
 \label{subsubsec:rec-dom-eq}
-% TODO consider new structure (the below)
-% First introduce c.o.f.e.s and the structures you can freely use to get a solution for your recursive domain equation
-% Then introduce all the specialized gunk necessary for our worlds
-%
-%solving the recursive domain equation means that we can present the actual definitions
-% Also see the notes from Aina
 In the previous sections, we sketched what our worlds should be.
-However, the worlds we want constitute a self-referential domain equation for which no solution exists in set theory.
+However, the worlds we want constitute a self-referential domain equation for which no solution exists in set and domain theory.
 % Move to a different domain
 Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
-\dominique{mention that this is a standard technique (cite the relevant papers) and an abstracted form of step-indexing (cite some more papers).}
-We move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
-Intuitively, a c.o.f.e.\ is a set with some extra structure.
+Solutions to recursive domain equations can be found using standard techniques~\citep{scott_1976,america_1989,Birkedal:2011:SKM:1926385.1926401}.
+Essentially, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
+A c.o.f.e.\ can be thought of as a set with added structure.
 Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
-In this section we present some of the technical details on solving as well as the final world domain equation (Theorem\ref{thm:rec-dom-eq-sol}).
-This section may be skipped by readers with no interest in the recursive domain equation.
+In this section, we present some of the technical details on solving our world equation as well as the final world equation (Theorem\ref{thm:rec-dom-eq-sol}).
+This section does not provide further intuition about the worlds and can be skipped.
 
@@
 \dominique{Introduce phrases like ``up to n steps'' etc.}
-% TODO: Lau write intuition for sequences and why they are necessary.
 \begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
@@
 \end{definition}
-% TODO write something here about why the different things (the meassure of the ofe and the guaranteed limits of the cofe) gives us in terms of finding a solution.
 In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
@@
 The later c.o.f.e.\ is used to make sure that the index decreases before the self-reference which gives something to recurse on in order to find a solution for the domain equation.
-If we only use the standard constructions to construct a self-referential equation and make sure to guard every negative occurrence of the self reference with the later c.o.f.e.\ construction, then the equation has a solution.
-\todo[inline]{Insert the solution theorem here?}
+If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.
 
@@
 \end{theorem}
-% TODO informally describe how solution to recursive domain equation is found (in terms of what we have to show, not the fixed-point theorem).
 %% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
 %% Aleš note? Tutorial note?
-The solution to the recursive domain equation is the preordered c.o.f.e.\ $\Wor$ in Theorem~\ref{thm:rec-dom-eq-sol}.
+Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} to construct the solution $\Wor$ to the recursive equation.
 Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
@@
 % TODO look over the original POPL paper to make sure to reuse all the text that makes sense to reuse.
-% TODO add an introduction to c.o.f.e.'s that describe how standard c.o.f.e. constructions always can be used to construct something with a solution.
 \end{jversion}

### 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Fix some notation

@@
                        \end{cases}\\
-    \purePart{W_\var{priv}} &\defeq \lambda r\ldotp
+    \purePart{W_\var{call\_stk}} &\defeq \lambda r\ldotp
                        \begin{cases}
-                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{priv}(r) = ((\spatialo,\var{sm}),\opc)\\
-                         W_\var{priv}(r) & \text{otherwise}
+                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{call\_stk}(r) = ((\spatialo,\var{sm}),\opc)\\
+                         W_\var{call\_stk}(r) & \text{otherwise}
                        \end{cases}\\
@@
     \array{l}
-    \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\
+    \text{for $i \in \{\mathrm{heap},\mathrm{free\_stk},\mathrm{call\_stk} \}$} \\
     \quad\exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
@@
 \[
-  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \defeq \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \; \hat{W} \defeq \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
     \begin{array}{l}
       \dom(\ms_S) = \dom(\ms_T) \wedge \\
-      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(W) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
+      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(\hat{W}
+      ) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
       \quad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
@@
         \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \\
-        \quad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spatialo,\square}, \aaddr'+\calllen) \tand \\
+        \quad\pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms_S,\ms_T)]{\spatialo,\square} \; (\ta,\stkb), \aaddr'+\calllen) \tand \\
         \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
@@
 The fact that it is static signifies that the contents will remain the same.
-The region that governs the stack frame must come from the private world which means that it is paired with a return address.
+The region that governs the stack frame must come from the call-stack sub-world which means that it is paired with a return address.
 The return address should correspond to an actual return address of a call in $\code$.

### 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Fix citation

@@
 \end{theorem}
-%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
-%% Aleš note? Tutorial note?
-Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} to construct the solution $\Wor$ to the recursive equation.
+Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,birkedal_taste_2014} to construct the solution $\Wor$ to the recursive equation.
 Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.

### 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Update figure

@@
 
-  \draw[gray, fill=gray, fill opacity=.4] (3,5) rectangle (5,0.5) node[fitting node] (w2l) { };
+  \draw[darkgray, fill=darkgray, fill opacity=.4] (3,5) rectangle (5,0.5) node[fitting node] (w2l) { };
   \draw (w2r2) edge[->,in=0, out=210] (w2l.east);
@@
 
-  \node[draw,rectangle, black, minimum size=2mm] (w1r) at (-2,5) {  };
+  \node[draw,rectangle, black, minimum size=4mm] (w1r) at (-2,5) {  };
   \node[fit=(w1r),black,inner sep=-\pgflinewidth,cross out, draw] (test) { };
@@
 
-  \node[draw,rectangle, black, minimum size=2mm] (w2r) at (6.5,5) {  };
+  \node[draw,rectangle, black, minimum size=4mm] (w2r) at (6.5,5) {  };
   \node[fit=(w2r),black,inner sep=-\pgflinewidth,cross out, draw] (test) { };
@@
 
-  \draw[gray, fill=gray, fill opacity=.4] (3,5) rectangle (5,0.5) node[fitting node] (w2l) { };
+  \draw[lightgray, fill=lightgray, fill opacity=.4] (3,5) rectangle (5,0.5) node[fitting node] (w2l) { };
   \draw (w2r3) edge[->,in=0, out=210] (w2l.east);
@@
     &\node[red, fill=red, fill opacity=.5,minimum size=0.4cm,label=right:{\footnotesize Shared region}] {}; \\
-    &\node[gray, fill=gray, fill opacity=.4,minimum size=0.4cm,label=right:{\footnotesize Shadow region}] {}; \\
+     \node[darkgray, fill=darkgray, fill opacity=.4,minimum size=0.4cm] {};
+    &\node[lightgray, fill=lightgray, fill opacity=.4,minimum size=0.4cm,label=right:{\footnotesize Shadow regions}] {}; \\
      \node[green, fill=green, fill opacity=.7,minimum size=0.4cm] {};
-    &\node[blue, fill=blue, fill opacity=.6,minimum size=0.4cm,draw,label=right:{\footnotesize Spatial region}] {}; \\
+    &\node[blue, fill=blue, fill opacity=.6,minimum size=0.4cm,draw,label=right:{\footnotesize Spatial regions}] {}; \\
     &\node[draw,rectangle, black, minimum size=4mm] (rev) {};

### 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add untrusted value relation mono lemma

@@
 
-\lau{TODO: insert monotonicity lemma}
+Words given to untrusted code may be stored by the untrusted code for later use.
+This means that a word is only safe to give to untrusted code if it is safe for them to keep it.
+This means that the untrusted value relation must be monotone with respect to the world, so words stay safe in any possible future world.
+\begin{lemma}[Untrusted value relation monotonicity]
+  \label{lem:monotonicity}
+  For all integers $n$, words $w_1$ and $w_2$, and worlds $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$.
+\end{lemma}
+
 \paragraph{The trusted value relation}

### 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Update text about mono lemma

--- 2019-07-31 popl-2019-journal-v/proving-full-abstraction.tex
@@
 
-\lau{TODO: insert monotonicity lemma}
+When we give a word to untrusted code, we can make no assumptions on when they will use it.
+For instance, they may store it in memory and use it in a later call.
+This means that a safe word must not only be safe now but also at any point in the future.
+The untrusted value relation ensures this as it is monotone with respect to future worlds.
+\begin{lemma}[Untrusted value relation monotonicity]
+  \label{lem:monotonicity}
+  For all integers $n$, words $w_1$ and $w_2$, and worlds $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$.
+\end{lemma}
+
 \paragraph{The trusted value relation}

### 2019-08-01 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Correstions in solving rec dom eq section

@@
 The three kinds of memory are modelled by having three sub-worlds where each sub-world is its own little world in a traditional sense.
-Linearity is modelled by adding ownership to certain parts of the world, so capabilities can take ownership of part of the memory thus ensuring that they are the sole reference.
-The call stack is modelled by the memory satisfaction (the relation that decides whether two memories are related in the world) by ensuring that the memory is actually shaped like a stack.
+Linearity is modelled by adding ownership to certain parts of the world that capabilities can take ownership of ensuring that they are the sole reference for that part of memory.
+Memory satisfaction (the relation that decides whether two memories are related in the world) models the call stack by ensuring that the memory is actually shaped like a stack.
 Finally, the seals are modelled by seal invariants that make sure that seals are only used on permitted sealables.
@@
 \]
-The sub-worlds are partial maps from names $\RegName$, modelled as natural numbers, to invariants.
+The sub-worlds are partial maps from names $\RegName$ (not to be confused with register names), modelled as natural numbers, to invariants.
 In order to define what this actually means, we need to be more precise about what we mean by invariants.
@@
 A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
-Intuitively, we want to be able to say that two memory segments are related when their contents are related.
+Intuitively, we want to be able to say that two memory segments are related when their content is related.
 That is, for every address in the two memory segments, the words that reside there must be related.
@@
 We cannot let multiple linear capabilities reference the same memory.
-The spatial region represent ownership a part of memory which ensures that a linear capability is not aliased.
+The spatial region represent ownership of a part of memory which ensures that a linear capability is not aliased.
 Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
@@
 The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
-The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities uniquely references part of memory.
+The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities \emph{uniquely} references part of memory.
 
@@
                        \begin{cases}
-                         (\spatial,\var{sm}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{sm})\\
+                         (\spatial,\var{H}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{H})\\
                          W_{\var{heap}}(r) & \text{otherwise}
@@
                        \begin{cases}
-                         ((\spatial,\var{sm}),\opc) & \text{if } W_\var{call\_stk}(r) = ((\spatialo,\var{sm}),\opc)\\
+                         ((\spatial,\var{H}),\opc) & \text{if } W_\var{call\_stk}(r) = ((\spatialo,\var{H}),\opc)\\
                          W_\var{call\_stk}(r) & \text{otherwise}
@@
                        \begin{cases}
-                         (\spatial,\var{sm}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{sm})\\
+                         (\spatial,\var{Hs}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{Hs})\\
                          W_{\var{free}}(r) & \text{otherwise}
@@
 Specifically, \stktokens{} has certain assumption on the seals used for return capabilities and closures.
-For instance, return seal must only be used to seal the return pointer of one specific return point.
+For instance, a return seal must only be used to seal the return pointer of one specific return point.
 Therefore, in addition to a relation on memory segments, some regions also carry a \emph{seal interpretation function} that relates the sealables that may be sealed with a given seal.
@@
 In Kripke logical relations, this kind of assumption changes, that do not invalidate the relatedness of values, are modelled by the future world relation $\future$.
-The future world relation can also be thought of as the model of memory changes over time.
+The future world relation can also be thought of as the model of changes in memory over time.
 
@@
 \]
-We make a similar change to $\Regionh$, and we make the seal invariant monotone as well.
-
-Kripke future world relation must be defined such that future worlds are extensions of the past world.
+We make a similar change to $\Regionh$. The seal invariants are monotone as well.
+
+Kripke future world relations must be defined such that future worlds are extensions of the past world.
 In other words, future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
 However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on assumption anymore.
-Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it sure to repurpose the memory.
+Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory.
 In terms of the world, this corresponds to dropping a region and replacing it with a new one.
@@
 This may be confusing as a $\spatial$-region does not mean that you ``have'' the capability, but we just explained that having a linear capability means that you can repurpose the memory it points to.
-As we will see later, the logical relation makes disjoint splits of ownership, so each capability gets its own ownership to depend on.
-A linear capability that is repurposed will get an entirely new world to depend on.
-After all, a future world would have to respect the old assumptions, but a repurposed capability needs to depend on new assumptions.
-The new world constructed for the repurposed linear capability would have a revoked region in place of the $\spatialo$-region, but the new world would not be a future world of the world it depended on before.
-While the repurposed capability gets a new world to depend on, all other capabilities in the system must remain valid in a future world to the one they depend on.
+As we will see later, the logical relation splits worlds with respect to their ownership, so each capability can get its own world with unique ownership to depend on.
+A repurposed linear capability gets an entirely new world to depend on.
+The new world constructed for the repurposed linear capability would have a revoked region in place of the $\spatialo$-region, but the new world would not be a future world of the world the linear capability depended on before.
+After all, a future world has to respect old invariants, but a repurposed capability needs to depend on new invariants.
+While the repurposed capability gets a new world to depend on, all other capabilities in the system depend on the same invariants, so they must be valid with respect to worlds that respect the old invariants, i.e.
+future worlds.
 Further, the future worlds must be compatible (agree on everything but ownership) with the new world.
@@
 Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes the linear capabilities affine.
-The world a capability is safe with respect to is no longer needed when the capability is dropped, so other worlds may pick up the ownership which the future region relation allows.
+A linear capability is safe with respect to a world with some $\spatialo$ regions.
+However, this world, and thus the $\spatialo$ regions, is no longer needed if the linear capability is dropped, so other worlds are free to claim the $\spatialo$ region in place of their $\spatial$ region.
 % describe why "Spatial -> spatialo" ~ affine or at least why we have it?
@@
 \end{definition}
-
-We cannot rely entirely on the standard constructions to create the worlds we need.
-Instead, we need to define c.o.f.e.'s and preordered c.o.f.e.'s that meet our needs.
-
 Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
@@
 \end{definition}
-The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step and up and requires the rest to be equal.
+The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
 This means that everything is equal at step $0$ as everything is left out.

### 2019-08-05 popl-2019-journal-v/discussion.tex
Minor modification and remark

@@
 Ideally, it should just combine security proofs for the compiler's other functionality with our results about \stktokens{}.
-We want to point out that our formulation enables such reuse.
+We point out that our formulation enables such reuse.
 Specifically, the compiler could be factored into a part that targets \srccm{}, followed by our embedding into \trgcm{}.
@@
   In order for the continuations and exceptions to work with the stack, they would have to be sealed with the return seals.
+  \DDin{I'm not convinced that these limitations cannot be avoided. Perhaps we should keep this discussion more tentative?}
   This requirement puts two strong limitations on continuations and exceptions.

### 2019-08-05 popl-2019-journal-v/discussion.tex
[POPLjv] Rewrite vertical composition discussion

@@
 
-\begin{thesisonly}
-  % Say something about horizontal vs vertical composition.
-  Fully-abstract compilers compose vertically which means that if we defined a compiler phase that targets \srccm{}, then we could use the full-abstraction theorem immediately.
-  However, fully-abstract compilers do not extend horizontally in the sense that if we add a new feature in the target and source language, then the full-abstraction theorem does not follow immediately.
-  Intuitively, this is the case because the new feature may have an unforeseen interaction with existing features that breaks a desired property.
-  In order to enforce more language properties on capability machines, it is reasonable to assume that new capability primitives may be necessary.
-  \stktokens{}, for instance, requires linear capabilities, and \citet{skorstengaard_reasoning_2017} uses local capabilities to enforce WBCF and LSE.
-  In order to implement address-hiding (i.e.\ have references that do not expose the address they actually point to), it is reasonable to believe that one would need a new capability primitive.
-  After all, the capabilities we use here do not hide the addresses in any meaningful way, so a new ``address hiding'' capability would be necessary to enforce this property.
-  If we extended \trgcm{} and \srccm{} with such new capabilities, then we would have to redo the entire full-abstraction proof.
-\end{thesisonly}
+  % Say something about full-abstractions proofs and feature extensions
+  Full-abstraction is a property of the whole language.
+  In other words, a full-abstraction proof must consider all features of a language to make sure that the features don't interact in a way that breaks language abstractions.
+  If we have a fully abstract compiler and add a feature to the source and target language, then the new compiler is no necessarily fully-abstract.
+  Full abstraction would have to be proven for the new compiler to make sure that the new feature does not break existing abstractions in the language.
+  Our proof of full-abstraction for \stktokens{} targets a simple capability machine that may not be able to enforce the high-level language abstractions we want, e.g.\ address hiding.
+  In other words, the full-abstraction proof cannot be reused immediately.
+  However, \stktokens{} is still a good candidate for the enforcement mechanism for well-bracketed control-flow and local-state encapsulation in a real fully abstract compiler.
+  Generally speaking, it is worth investigating enforcement mechanisms for full abstraction in a simple setting that allows to quickly try out ideas and verify that the enforcement works.
 

### 2019-08-05 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Moderate wording about revoked regions

@@
 Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory.
-In terms of the world, this corresponds to dropping a region and replacing it with a new one.
-Due to the fact that future worlds must be an extension of the past world, we cannot drop a region.
-Instead, we mark it as revoked which for all intends and purposes corresponds to dropping the region.
+We mark dropped regions as revoked in the world following \citet{ahmed_2004} and \citet{thamsborg_2011} which for all intends and purposes corresponds to actually dropping the region.
 

### 2019-08-05 popl-2019-journal-v/discussion.tex
[POPLjv] Moderate wording about revoked regions
[POPLjv] Rewrite stack allocation discussion

@@
   In order for the continuations and exceptions to work with the stack, they would have to be sealed with the return seals.
-  \DDin{I'm not convinced that these limitations cannot be avoided. Perhaps we should keep this discussion more tentative?}
+  \dominique{I'm not convinced that these limitations cannot be avoided. Perhaps we should keep this discussion more tentative?}
   This requirement puts two strong limitations on continuations and exceptions.
@@
 \begin{jversion}
-Programs that use a C-like calling convention often allow programs to pass a stack reference in a call either to provide a reference for a piece of memory to work on or to indicate where the result of the call should be placed.
-\stktokens{} does not support this as it expects to get a stack pointer for the entire stack it handed out.
-\dominique{Err... I don't see the problem.
-  Passing stack references to callees should work fine, as long as the callee treats them linearly (i.e.\ doesn't try to copy them).
-  }
-However, if we have allocated part of the stack for a different purpose and handed out a reference for it, then the stack pointer cannot reference the same memory as the stack pointer is a linear capability.
-\dominique{I don't understand this sentence.}
-However, while it is not possible to allocate a reference for the stack, one can easily imagine that the stack could still be used to pass and return values.
-\dominique{Or this one...}
+  Programming languages with a C-like calling convention often allow programs to pass stack references in calls.
+  \stkotkens{} supports stack references but with a couple of caveats.
+  First of all, the stack capability is linear, so all references to the stack have to be linear.
+  This means that the callee has to treat references linearly.
+  Next like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (which allows them to return). 
+  Finally, the encapsulated local stack frame should be a continuous piece of memory which means that the stack allocations need to be next to the stack frame.
+  An escape analysis could be used to statically determine where to put allocations and, in principle, the allocations could be reordered dynamically before a call.
+  It is not clear to us to what extend existing C-code is covered by the stack references supported by \stktokens{}.
 \end{jversion}

### 2019-08-05 popl-2019-journal-v/discussion.tex
[POPLjv] Rewrite continuation/exception discussion

@@
 
-  We conjecture that \stktokens{} can support a limited form of continuations and exceptions.
-  Specifically, the continuations and exceptions would be passed to the callee as sealed code capabilities.
-  In order for the continuations and exceptions to work with the stack, they would have to be sealed with the return seals.
-  \dominique{I'm not convinced that these limitations cannot be avoided. Perhaps we should keep this discussion more tentative?}
-  This requirement puts two strong limitations on continuations and exceptions.
-  First of all, they can be used by the immediate callee, but they cannot be used in subsequent calls made by the callee (assuming they use \stktokens{}).
-  The problem is that the callee would have to take up part of the stack for their local stack frame.
-  This would break up the return token that should be used with the continuation (or exception).
-  If the callee really needs to pass on the continuation, then they would have to construct a wrapper that unravels the stack and restores the necessary return token.
-  Second, a call that passes continuations or exceptions in this way has to pass the same continuations and exceptions in every call.
-  The problem is that just like the return code capability, the continuation and exception capabilities are non-linear.
-  This means that a callee can store them for subsequent calls preventing us from using different continuations and exceptions for different calls.
-
-  \dominique{Also: should we mention the limited additional control flow we seem to get if we omit the stack base check (where  adversaries can use a kind of non-preemptive multi-threading)?}
+  We have not investigated support for continuations and exceptions in \stktokens{} thoroughly but believe they are supported.
+  The continuations and exceptions could plausibly be implemented similarly to return pointers, i.e.\ as sealed code capabilities.
+  The exception capability would point to a piece of code handling the specific exception, and one would cast an exception by invoking the exception capability.
+  The continuations and exceptions would be sealed with the return capability allowing them to be used with the appropriate encapsulated stack frame.
+  The stack is crucial to \stktokens{}, and it must be unravelled as an exception falls through calls.
+  One way to handle stack unravelling is to make the code responsible for unravelling its own stack frame.
+  For a piece of code that wants to pass on a continuation, it would have to pass on a capability for a piece of code that unravels the stack and uses the continuation.
+  Alternatively, it may be possible to have a piece of central trusted code that does the stack unravelling.
 

### 2019-08-05 popl-2019-journal-v/discussion.tex
[POPLjv] Rephrase sentence

@@
   One way to handle stack unravelling is to make the code responsible for unravelling its own stack frame.
-  For a piece of code that wants to pass on a continuation, it would have to pass on a capability for a piece of code that unravels the stack and uses the continuation.
+  To pass on a continuation, it would have to be wrapped in some sort of continuation forward record that unravels the stack and calls the continuation.
   Alternatively, it may be possible to have a piece of central trusted code that does the stack unravelling.

### 2019-08-05 popl-2019-journal-v/paper.tex
[POPLjv] Small changes
[POPLjv] Fix some notation
[POPLjv] Write paragraph for Stelios related work

@@
 
+\renewcommand{\pwpriv}[1][W]{#1.\mathrm{call\_stk}}
+\renewcommand{\pwfree}[1][W]{#1.\mathrm{free\_stk}}
 \renewcommand{\lrr}{\lrrg{ }}
@@
 
-%TODO distinguish closure seals from return seals.
 The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
@@
 Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed with.
-
-% TODO Mention sealed capabilities?
 
@@
 % Full abstraction thm
-% \todo[inline]{Look at coloring in these theorems. Why are all components and contexts blue?}
 \begin{theorem}
@@
 Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
-% \todo[inline]{Consider writing a bit more about compiler correctness and contextual equivalence preservation.}
 
@@
 
+Local capabilities can be used to ensure well-bracketed control-flow and local-state encapsulation as demonstrated by \citet{skorstengaard_reasoning_2017}.
+Recently, \citet{tsampas_2019} demonstrated that an extension of local capabilities to multiple linearly ordered colours can be used to enforce the life time of stack references.
+Specifically, a stack references should not be able to outlive the stack frame it points to.
+If \stktokens{} was extended with stack references, then it would also enforce reference life times.
+Specifically in order to return from a call, we must use the return token, i.e.\ the stack.
+The stack is linear, so if there are references to it, aside from the stack capability itself, then we cannot have a complete return token.
+This means that we have to splice all the stack references together with the stack capability to complete the return token in order to return.
+\citet{tsampas_2019} allow (almost) normal references that can be stored in multiple places at the same time.
+This means that their approach is more like C than what \stktokens{} has to offer.
+
 \dominique{TODO Relate full-abstraction to other notions of secure compilation}
-
-\dominique{TODO Relate to Stelios' CSF 2019 paper}
 

### 2019-08-05 popl-2019-journal-v/paper.tex
[POPLjv] Various typesetting
Related work on linear capabilities.

@@
 \section{Related Work}
-In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation.
+In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation or the linear capabilities we use to do it.
 We do not repeat the work we discussed in Section~\ref{sec:introduction}.
@@
 
-\dominique{TODO Relate to Thomas Van Strydonck's ICFP19 paper which uses linear capabilities (at a higher level) to enforce separation logic in a fully abstract way.}
+In addition to the already-mentioned work on linear capabilities, \citet{van_strydonck_linear_2019} have recently used them in a secure (fully abstract) compiler for a C-like language with separation logic contracts.
+A form of linear capabilities was also used in the SAFE machine developed within the CRASH/SAFE project \citep{DBLP:conf/sp/AmorimDGHPST15,DBLP:journals/jcs/AmorimCDDHPPPT16}.
 

### 2019-08-05 popl-2019-journal-v/proving-full-abstraction.tex

@@
       \begin{array}{l}
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \vee\\
         (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
@@
       \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)} \wedge\\
+      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc} \wedge\\
       \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
@@
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \\
+      \qquad\memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W_\lrm} \\
       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
@@
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \wedge\\
-      \qquad\memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \wedge \\
+      \qquad\memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W_\lrm} \wedge \\
       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
@@
      &\npair{\var{comp},\var{comp}} \;\mid \;\\
-      &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \tand \\
+      &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \text{ and} \\
       &\qquad\text{For all } W' \future W \ldotp \\
@@
       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \tand\\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \text{ and }\\
       &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
@@
      \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
-     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \tand\\
+     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \text{ and}\\
      \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}

### 2019-08-06 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add colour to definitions

@@
 \[
-  \npair{(\overline{\sigma},ms,\ms_T)} \in \lrheap(\pwheap)(W') = 
+  \npair{(\overline{\sigma},\src{ms},\ms_T)} \in \lrheap(\pwheap)(W') = 
   \left\{
@@
       \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
-      \quad \ms = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \src{\ms} = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
       \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
@@
 \[
-  \memSatFStack{ms_\stk,\ms_T}{W} \text{ iff } 
+  \memSatFStack{\src{\ms_\stk},\ms_T}{W} \text{ iff } 
   \left\{
@@
       \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_2(R_\ms(r)) \wedge \\
-      \quad \ms_\stk = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
-      \quad \stkb \in \dom(\ms_T) \wedge \stkb \in \dom(\ms_\stk) \wedge \\
+      \quad \src{\ms_\stk} = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad \stkb \in \dom(\ms_T) \wedge \stkb \in \dom(\src{\ms_\stk}) \wedge \\
       \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World\ldotp\\
@@
 \[
-  \memSatStack{\stk,\ms_T}{W} \text{ iff } 
+  \memSatStack{\src{\stk},\ms_T}{W} \text{ iff } 
   \left\{
     \begin{array}{l}
-      \gc = (\_,\stkb,\_,\_) \wedge \\
-      W_\var{stack} = \pwpriv \wedge \\
-      \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
-      \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
-      \quad \forall j < i \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
-      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
-      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
-      \quad \ms_0 \uplus \dots \uplus \ms_m = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
-      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World \ldotp \\
-      \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
-      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n\ldotp \\
-      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
-      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
-      \qquad \quad \exists i \ldotp \opc_i = W_\var{stack}(r).\opc \wedge \ms_i = \pi_1 (R_\ms(r))
+      \exists m, \src{\opc_0},\dots,\src{\opc_m}, \src{\ms_0}, \dots, \src{\ms_m}, W_\var{stack}\ldotp\\
+      \quad\gc = (\_,\stkb,\_,\_) \wedge \\
+      \quad W_\var{stack} = \pwpriv \wedge \\
+      \quad\src{\stk} = \src{(\opc_0,\ms_0), \dots (\opc_m,\ms_m)} \wedge \\
+      \quad\forall i \in \{0,\dots,m\} \ldotp (\dom(\src{\ms_i}) \neq \emptyset \wedge\\
+      \quad\quad \forall j < i \ldotp \forall a \in \dom(\src{\ms_i}) \ldotp \forall a' \in \dom(\src{\ms_j}) \ldotp \stkb < a < a') \wedge\\
+      \quad\exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
+      \quad\quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
+      \quad\quad \src{\ms_0} \uplus \dots \uplus \src{\ms_m} = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
+      \quad\quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World \ldotp \\
+      \quad\qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
+      \quad\qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n\ldotp \\
+      \quad\qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+      \quad\qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
+      \quad\qquad \quad \exists i \ldotp \src{\opc_i} = W_\var{stack}(r).\opc \wedge \src{\ms_i} = \pi_1 (R_\ms(r))
     \end{array}
@@
 \[
-  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
+  \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff } 
   \left\{
     \begin{array}{l}
-      \exists \opc_0, \dots \opc_m, \ms_0, \dots, \ms_m, W_{\var{stack}}, W_{\var{free\_stack}}, W_{\var{heap}} \ldotp\\
-      \quad \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
-      \quad \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \text{ is defined} \wedge\\
+      \exists m,  \src{\opc_0}, \dots \src{\opc_m}, \src{\ms_0}, \dots, \src{\ms_m}, W_{\var{stack}}, W_{\var{free\_stack}}, W_{\var{heap}} \ldotp\\
+      \quad \src{\stk} = \src{(\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m)} \wedge \\
+      \quad \src{\ms_S} \# \src{\ms_\stk} \# \src{\ms_0} \# \dots \# \src{\ms_m} \wedge\\
       \quad W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
-      \quad\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}\ldotp \\
-      \qquad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
+      \quad\exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \src{\ms_{S,f}}, \src{\ms_S'},\overline{\sigma}\ldotp \\
+      \qquad \src{\ms_S} = \src{\ms_{f,S}} \uplus \src{\ms_S'} \wedge \\
       \qquad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
@@
       \qquad \baddr_\stk -1,\eaddr_\stk + 1 \in \dom(\ms_{T,f}) \wedge \\
-      \qquad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
-      \qquad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
-      \qquad \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
+      \qquad \memSatStack{\src{\stk},\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
+      \qquad \memSatFStack{\src{\ms_\stk},\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
+      \qquad \npair{(\overline{\sigma},\src{\ms_S'},\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
     \end{array}
@@
   \hline
-  \npair{(w_S,w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
-  \npair{(w_S,w_T)} & \lrvtrusted(W) & values (machine words) & \\
-  \npair{(\reg_S,\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
-  \npair{\Phi_S,\Phi_T}  &  \lro & execution configurations & \\
-  \npair{(w_S,w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
-  \left(\arraycolsep=1pt\array{l}(w_{S,1},w_{S,2}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
-  \multicolumn{2}{|c|}{$\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
+  \npair{(\src{w_S},w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
+  \npair{(\src{w_S},w_T)} & \lrvtrusted(W) & values (machine words) & \\
+  \npair{(\src{\reg_S},\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
+  \npair{\src{\Phi_S},\Phi_T}  &  \lro & execution configurations & \\
+  \npair{(\src{w_S},w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
+  \left(\arraycolsep=1pt\array{l}(\src{w_{S,1}},\src{w_{S,2}}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
+  \multicolumn{2}{|c|}{$\memSat{\src{\ms_S},\src{\stk},\src{\ms_\stk},\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
   \hline
@@
 \[
-  H_A^{\mathrm{std},\square} \; \gc \; \hat{W} \defeq \left\{ \npair{\ms_S,\ms_T} \middle|
-    \begin{array}{l}
-      \dom(\ms_S) = \dom(\ms_T) = A \wedge \\
+  H_A^{\mathrm{std},\square} \; \gc \; \hat{W} \defeq \left\{ \npair{\src{\ms_S},\ms_T} \middle|
+    \begin{array}{l}
+      \dom(\src{\ms_S}) = \dom(\ms_T) = A \wedge \\
 % There should be a later on the right side of the equality? Lau: To myself, on second thought we leave that implicit. This should be fine.
       \exists S : A \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in A} S(\aaddr) \wedge\\ 
-      \quad \forall \aaddr \in A \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \quad \forall \aaddr \in A \ldotp \npair{(\src{\ms_S}(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
     \begin{array}{l}
-      \dom(\ms_S) = \dom(\ms_T) \wedge \\
-      \exists S : \dom(\ms_S) \fun \World \ldotp \xi(\hat{W}
-      ) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
-      \quad \forall \aaddr \in \dom(\ms_S) \ldotp \npair{(\ms_S(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
+      \dom(\src{\ms_S}) = \dom(\ms_T) \wedge \\
+      \exists S : \dom(\src{\ms_S}) \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
+      \quad \forall \aaddr \in \dom(\src{\ms_S}) \ldotp \npair{(\src{\ms_S}(\aaddr),\ms_T(\aaddr))} \in \lrv(S(\aaddr))
     \end{array}
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\src{\retptrc(\baddr,\eaddr,\aaddr'+\calllen)},\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
       \begin{array}{l}
@@
     \begin{array}{l}
-\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
+\left. \npair{\arraycolsep=0pt\left(\array{l}\src{\retptrd(\baddr,\eaddr)},\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
       \begin{array}{l}
@@
     \begin{array}{l}
-\left. \npair{(\vsc, \vsc' )} \middle| \right. \\
-      \begin{array}{l}
-        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \vee\\
+\left. \npair{(\src{\vsc}, \vsc' )} \middle| \right. \\
+      \begin{array}{l}
+        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\src{\vsc},\vsc')} \in \lrv \; \xi(\hat{W})) \vee\\
         (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \tand \sigrets \subseteq \gsigrets \tand \\
-         \quad((\exec{\vsc} \wedge \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
-         \quad\ (\nonExec{\vsc} \wedge\npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W}))))
+         \quad((\exec{\src{\vsc}} \wedge \npair{(\src{\vsc},\vsc')} \in \lrvtrusted \; \xi(\hat{W})) \vee\\
+         \quad\ (\nonExec{\src{\vsc}} \wedge\npair{(\src{\vsc},\vsc')} \in \lrv \; \xi(\hat{W}))))
       \end{array}
@@
   \label{lem:monotonicity}
-  For all integers $n$, words $w_1$ and $w_2$, and worlds $W' \future W$, if $\npair{(w_1,w_2)} \in \lrv(W)$, then $\npair{(w_1,w_2)} \in \lrv(W')$.
+  For all integers $n$, words $\src{w_1}$ and $w_2$, and worlds $W' \future W$, if $\npair{(\src{w_1},w_2)} \in \lrv(W)$, then $\npair{(\src{w_1},w_2)} \in \lrv(W')$.
 \end{lemma}
@@
       \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
-      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrv(S(r))
+      \qquad\npair{(\src{\reg_S}(r),\reg_T(r))} \in \lrv(S(r))
     \end{array}
@@
       \qquad\memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W_\lrm} \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
-      \qquad \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}}\\
+      \qquad\src{\Phi_S} = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\\
+      \qquad\src{\Phi_S'} = \src{\Phi_S} \updReg{\pcreg}{\src{v_{c,S}}}\\
       \qquad\Phi_T = (\ms_T,\reg_T)\\
@@
       \qquad W \oplus W_\lrrs \oplus W_\lrm\\
-      \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+      \qquad\Rightarrow\npair[n']{\left(\src{\Phi_S'}, \Phi_T' \right)}\in \lro
     \end{array}
@@
       \qquad\memSat[n']{\src{\ms_S,\stk,\ms_\stk},\ms_T}{W_\lrm} \wedge \\
-      \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
+      \qquad\src{\Phi_S} = \src{(\ms_S,\reg_S,\stk, \ms_\stk)}\wedge\\
       \qquad\Phi_T = (\ms_T,\reg_T) \wedge\\
       \qquad W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
-      \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp\\
-      \quad\qquad \Phi_S' = \xjumpResult{v_{c,S}}{v_{d,S}}{\Phi_S} \tand\\
+      \qquad\Rightarrow \exists \src{\Phi_S'},\Phi_T'\ldotp\\
+      \quad\qquad \src{\Phi_S'} = \xjumpResult{\src{v_{c,S}}}{\src{v_{d,S}}}{\src{\Phi_S}} \tand\\
       \quad\qquad\Phi_T' = \xjumpResult{v_{c,T}}{v_{d,T}}{\Phi_T}\tand\\
-      \quad\qquad\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
+      \quad\qquad\npair[n']{\left(\src{\Phi_S'}, \Phi_T' \right)}\in \lro
     \end{array}
@@
   \left\{\begin{aligned}
-     &\npair{\var{comp},\var{comp}} \;\mid \;\\
+     &\npair{\src{\var{comp}},\src{\var{comp}}} \;\mid \;\\
       &\qquad\var{comp} = (\mscode,\msdata,\overline{a_{\mathrm{import}} \mapsfrom s_{\mathrm{import}}},\overline{s_{\mathrm{export}} \mapsto w_{\mathrm{export}}},\sigrets,\sigcloss) \text{ and} \\
       &\qquad\text{For all } W' \future W \ldotp \\
-      &\qquad\quad\text{If } \overline{\npair[n']{(w_{\mathrm{import}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
+      &\qquad\quad\text{If } \overline{\npair[n']{(\src{w_{\mathrm{import}}},w_{\mathrm{import}})}} \in \lrv(\purePart{W'}) \text{ for all $n' < n$}\\
       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\mscode\uplus \msdata', \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \text{ and }\\
-      &\qquad\quad\overline{\npair{(w_{\mathrm{export}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\src{\mscode\uplus \msdata'}, \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \text{ and }\\
+      &\qquad\quad\overline{\npair{(\src{w_{\mathrm{export}}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
   \end{aligned}
@@
     \begin{multlined}
-     \npair{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}),(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
-     \npair{(\var{comp}_0,\var{comp}_0)} \in \lrcomp(W) \text{ and}\\
-     \{(\_ \mapsto c_{\mathrm{main},c}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
+     \npair{\src{(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})},(\var{comp}_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})} \;\mid \;\\
+     \npair{(\src{\var{comp}_0},\var{comp}_0)} \in \lrcomp(W) \text{ and}\\
+     \{(\_ \mapsto \src{c_{\mathrm{main},c}}),(\_ \mapsto c_{\mathrm{main},d})\} \subseteq \overline{w_{\mathrm{export}}}
        \end{multlined}
@@
 \npair{\left(
-        (\ms_S,\reg_S,\stk,\ms_\stk),
+        \src{(\ms_S,\reg_S,\stk,\ms_\stk)},
         (\ms_T,\reg_T)\right)} \; \mid \; \\
@@
       \exists W_M,W_R,W_\pcreg \ldotp W = W_M \oplus W_R \oplus W_\pcreg \tand\\
-      \npair{( (\reg_S(\pcreg),\reg_S(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
-      \reg_S(\pcreg) \neq \retptrc(\_) \wedge 
-      \reg_S(\rdata) \neq \retptrd(\_) \wedge \\
-      \nonExec{\reg_S(\rdata)} \wedge
+      \npair{( (\src{\reg_S}(\pcreg),\src{\reg_S}(\rdata)), (\reg_T(\pcreg),\reg_T(\rdata)) )} \in \lrexj(W_\pcreg) \tand \\
+      \src{\reg_S}(\pcreg) \neq \src{\retptrc(\_)} \wedge 
+      \src{\reg_S}(\rdata) \neq \src{\retptrd(\_)} \wedge \\
+      \nonExec{\src{\reg_S}(\rdata)} \wedge
       \nonExec{\reg_T(\rdata)}\\
-      \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W_M} \tand\\
+      \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W_M} \tand\\
       \npair{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\})(W_R)

### 2019-08-06 popl-2019-journal-v/proving-full-abstraction.tex
Some remarks
Typo

@@
   \]
+  \dominique{I suspect the above definition is wrong and should say}
+  \[
+    x \nequal[n]_{\blater} x' \text{ iff } \left\{
+      \begin{array}{l}
+        n = 0 \vee{} \\
+        x \nequal[n-1]_X x'
+      \end{array}
+    \right.
+  \]
 \end{definition}
 We will also need the canonical preordered c.o.f.e.\ for sum.
+\dominique{products?}
+\dominique{I wonder if it's necessary to show all these constructions in full detail?}
 \begin{definition}[Product preordered c.o.f.e.]
@@
 \end{definition}
-Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:shared-reg-p-cofe} define the preoreded c.o.f.e.'s necessary for the regions in our worlds.
+Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:shared-reg-p-cofe} define the preordered c.o.f.e.'s necessary for the regions in our worlds.
 

### 2019-08-06 popl-2019-journal-v/discussion.tex
[POPLjv] Typo
A pass over some new parts of the discussion and related work sections

@@
 
-  We have not investigated support for continuations and exceptions in \stktokens{} thoroughly but believe they are supported.
-  The continuations and exceptions could plausibly be implemented similarly to return pointers, i.e.\ as sealed code capabilities.
-  The exception capability would point to a piece of code handling the specific exception, and one would cast an exception by invoking the exception capability.
-  The continuations and exceptions would be sealed with the return capability allowing them to be used with the appropriate encapsulated stack frame.
-  The stack is crucial to \stktokens{}, and it must be unravelled as an exception falls through calls.
-  One way to handle stack unravelling is to make the code responsible for unravelling its own stack frame.
-  To pass on a continuation, it would have to be wrapped in some sort of continuation forward record that unravels the stack and calls the continuation.
-  Alternatively, it may be possible to have a piece of central trusted code that does the stack unravelling.
+  We have not investigated support for continuations and exceptions in \stktokens{} thoroughly but we expect such support could be added.
+  For exceptions, one approach would let callers provide callees with an additional capability for exceptional returns.
+  This second capability would be similar to the code part of the return capability pair and signed with the same seal.
+  The callee would be able to invoke it to signal that an exception has been thrown after which the caller's code would handle the exception, either by executing an exception handler or by unwinding its stack frame and passing the exception on to its own caller.
+  Essentially, this would mean every function would be made responsible for unwinding its own stack frame.
+  Continuations are more complicated but could perhaps be treated using similar ideas.
+  Alternatively, it might also be possible to have a piece of central trusted code that does the stack unravelling for all stack frames.
+  To do this, this trusted code would need to receive a copy of all return seals from the linker.
 
@@
   \dominique{This paragraph should perhaps also point out the link to the dynamic compromise paper?}
+  \dominique{Then again, we come back to this later in the related work, so perhaps not necessary to discuss this here?}
   Our full-abstraction theorem, Theorem~\ref{thm:full-abstraction}, is not pure full-abstraction as it requires the components to be reasonable and well-formed.
@@
 
+\dominique{Start new subsection here?}
+
 \begin{jversion}
   Programming languages with a C-like calling convention often allow programs to pass stack references in calls.
-  \stkotkens{} supports stack references but with a couple of caveats.
+  \stktokens{} supports stack references but with a couple of caveats.
   First of all, the stack capability is linear, so all references to the stack have to be linear.
   This means that the callee has to treat references linearly.
-  Next like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (which allows them to return). 
-  Finally, the encapsulated local stack frame should be a continuous piece of memory which means that the stack allocations need to be next to the stack frame.
+  Next, like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (which allows them to return). 
+  Finally, the encapsulated local stack frame should be a contiguous piece of memory (because it has to be addressable by a single capability: the data part of the return capability pair).
+  Because of this, stack-allocated objects for which references are passed to callees must be allocated at the top or bottom of the caller's stack frame.
   An escape analysis could be used to statically determine where to put allocations and, in principle, the allocations could be reordered dynamically before a call.
-  It is not clear to us to what extend existing C-code is covered by the stack references supported by \stktokens{}.
+  In summary, support for passing stack references as arguments to callees could be added to \stktokens{}, but this would probably require some changes in the compiler and, more importantly, would require the callee to take special care when manipulating such references.
+  We are unsure whether it's realistic to apply this approach for existing C code.
 \end{jversion}
-
-\dominique{Perhaps worth mentioning the problem with adding support for linear values in a real C compiler: how to explain to the compiler that it should treat the linear values linearly and not use optimizations that will break this.}
 

### 2019-08-06 popl-2019-journal-v/paper.tex
A pass over some new parts of the discussion and related work sections

@@
 \newcommand{\erasen}[2]{\lfloor #1 \rfloor_{#2}}
+
+\renewcommand{\tand}{\wedge}
+\renewcommand{\tor}{\vee}
 
@@
 Local capabilities can be used to ensure well-bracketed control-flow and local-state encapsulation as demonstrated by \citet{skorstengaard_reasoning_2017}.
-Recently, \citet{tsampas_2019} demonstrated that an extension of local capabilities to multiple linearly ordered colours can be used to enforce the life time of stack references.
-Specifically, a stack references should not be able to outlive the stack frame it points to.
+Recently, \citet{tsampas_2019} demonstrated that an extension of local capabilities with multiple linearly ordered colours can be used to enforce the life time of stack references.
+Specifically, a stack reference should not be able to outlive the stack frame it points to.
 If \stktokens{} was extended with stack references, then it would also enforce reference life times.

### 2019-08-07 popl-2019-journal-v/paper.tex
[POPLjv] Add reference to micro-policy linear capabilities

@@
 A form of linear capabilities was also used in the SAFE machine developed within the CRASH/SAFE project \citep{DBLP:conf/sp/AmorimDGHPST15,DBLP:journals/jcs/AmorimCDDHPPPT16}.
+\citet{Abate:2018:GCG:3243734.3243745} used micro-policy enforced linear return capabilities to ensure a cross-component stack discipline.
+Their linear capabilities were designed specifically to enforce the stack discipline but behave similarly to ours with the notable difference that their linear return pointers are destroyed in a jump.
 

### 2019-08-07 popl-2019-journal-v/paper.tex
[POPLjv] Relate to other notions of secure compilation

@@
 
-\dominique{TODO Relate full-abstraction to other notions of secure compilation}
-
 In addition to the already-mentioned work on linear capabilities, \citet{van_strydonck_linear_2019} have recently used them in a secure (fully abstract) compiler for a C-like language with separation logic contracts.
@@
 Their linear capabilities were designed specifically to enforce the stack discipline but behave similarly to ours with the notable difference that their linear return pointers are destroyed in a jump.
+
+There are other notion of secure compilation than full-abstraction~\citep{abadi_protection_1998}.
+\citet{abate_2019} presents an overview of trace-based secure compilation properties.
+They show that full-abstraction is a weak notion of secure compilation relatively to other notions, e.g., Robust Relational Hyperproperty Preservation (RrHP) their strongest notion of secure compilation.
+However, other notions of secure compilation that promise WBCF and LSE must also enforce it somehow.
+\stktokens{} is an enforcement scheme that likely would work for stronger notions of secure compilation.
+In fact, we believe that our result could be lifted from full-abstraction to RrHP reusing the back-translation from the full-abstraction proof.
 

### 2019-08-07 popl-2019-journal-v/discussion.tex
[POPLjv] Minor textual changes

@@
   Alternatively, it might also be possible to have a piece of central trusted code that does the stack unravelling for all stack frames.
-  To do this, this trusted code would need to receive a copy of all return seals from the linker.
+  To do this, this the trusted code would need to receive a copy of all return seals from the linker.
 
@@
   Next, like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (which allows them to return). 
-  Finally, the encapsulated local stack frame should be a contiguous piece of memory (because it has to be addressable by a single capability: the data part of the return capability pair).
+  Finally, the encapsulated local stack frame should be a continuous piece of memory (because it has to be addressable by a single capability: the data part of the return capability pair).
   Because of this, stack-allocated objects for which references are passed to callees must be allocated at the top or bottom of the caller's stack frame.
@@
   In summary, support for passing stack references as arguments to callees could be added to \stktokens{}, but this would probably require some changes in the compiler and, more importantly, would require the callee to take special care when manipulating such references.
-  We are unsure whether it's realistic to apply this approach for existing C code.
+  We are unsure whether it is realistic to apply this approach for existing C code.
 \end{jversion}

### 2019-08-07 popl-2019-journal-v/appendix.tex
[POPLjv] First pass over solving the recursive domain equation section

@@
 
+  \section{Standard c.o.f.e.\ definitions}
+  \label{app:cofe}
+\begin{definition}[Product c.o.f.e.]
+\label{def:product-cofe}
+Given two c.o.f.e.'s $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal \right)_{n=0}^{\infty} \right)$ where the equivalence family is defined as for $(x,y),(x',y') \in X \times Y$
+\[
+  (x,y) \nequal (x',y') \text{ iff } x \nequal x' \wedge y \nequal y'
+\]
+\end{definition}
+\begin{definition}[Product preordered c.o.f.e.]
+  \label{def:product-p-cofe}
+  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
+  \[
+  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \]
+  where the preorder $\future$ distributes to the underlying preorder, i.e.\
+  \[
+\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
+  \]
+  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
+  \[
+\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
+  \]
+\end{definition}
+\begin{definition}[Union preordered c.o.f.e.]
+  \label{def:sum-p-cofe}
+  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
+  \[
+  \left(X \cup Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \]
+  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
+  \[
+    \text{for } z,z' \in X \cup Y,\quad  z' \future z \text{ iff } \left\{
+      \begin{array}{l}
+        z,z' \in X \wedge z' \future_x z \vee{} \\
+        z,z' \in Y \wedge z' \future_Y z
+      \end{array}
+    \right.
+  \]
+  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
+  \[
+    \text{for } z,z' \in X \cup Y,\quad  z \nequal z' \text{ iff } \left\{
+      \begin{array}{l}
+        z,z' \in X \wedge z \nequal_x z' \vee{} \\
+        z,z' \in Y \wedge z \nequal_Y z'
+      \end{array}
+    \right.
+  \]
+\end{definition}
+\begin{definition}[$\blater$ preordered c.o.f.e.]
+  \label{def:later-p-cofe}
+  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
+  \[
+    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
+  \]
+  where
+  \[
+    x \nequal[n]_{\blater} x' \text{ iff } \left\{
+      \begin{array}{l}
+        n = 0 \vee{} \\
+        x \nequal
+        [n-1]_X x'
+      \end{array}
+    \right.
+  \]
+\end{definition}
+
 %%% Local Variables:

### 2019-08-07 popl-2019-journal-v/proving-full-abstraction.tex
[POLjv] Correct blater definition
[POPLjv] First pass over solving the recursive domain equation section

--- 2019-08-06 popl-2019-journal-v/proving-full-abstraction.tex
@@
 \noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
-There are a number of standard c.o.f.e.\ constructions for countably infinite sets, products, sums, and functions.
+There are a number of standard c.o.f.e.\ constructions for countably sets, products, sums, and functions.
 A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
-The other constructions are used to create c.o.f.e.'s from other c.o.f.e.'s.
-For instance the product construction is defined as follows:
-\begin{definition}[Product c.o.f.e.]
-\label{def:product-cofe}
-Given two c.o.f.e.'s $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal \right)_{n=0}^{\infty} \right)$ where the equivalence family is defined as for $(x,y),(x',y') \in X \times Y$
-\[
-  (x,y) \nequal (x',y') \text{ iff } x \nequal x' \wedge y \nequal y'
-\]
-\end{definition}
-The remaining standard constructions are defined similarly.
-Another important construction is the \emph{later} ($\blater$) c.o.f.e.\
-The later c.o.f.e.\ construction is based on one c.o.f.e.\ which it decreases the index of.
+The other standard constructions are defined naturally and compose new c.o.f.e.'s from existing c.o.f.e.'s.
+
+In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
+This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
+Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
+To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
 \begin{lemma}[$\blater$ c.o.f.e.]
@@
   \[
-    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
+    x \nequal[n]_{\blater} x' \text{ iff } \left\{
       \begin{array}{l}
         n = 0 \vee{}\\
-        x \nequal_X x'
+        x \nequal[n-1]_X x'
       \end{array}
@@
 \end{lemma}
-The later c.o.f.e.\ is used to make sure that the index decreases before the self-reference which gives something to recurse on in order to find a solution for the domain equation.
+The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
+Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
 If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.
@@
 The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
-There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
-We will need the standard construction for later, product, and sum, so we define them here.
-The reader may skip these definitions for now and refer back to them later.
-\begin{definition}[$\blater$ preordered c.o.f.e.]
-  \label{def:later-p-cofe}
-  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
-  \[
-    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
-  \]
-  where
-  \[
-    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
-      \begin{array}{l}
-        n = 0 \vee{} \\
-        x \nequal
-        [n]_X x'
-      \end{array}
-    \right.
-  \]
-\end{definition}
-We will also need the canonical preordered c.o.f.e.\ for sum.
-\begin{definition}[Product preordered c.o.f.e.]
-  \label{def:product-p-cofe}
-  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
-  \[
-  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
-  \]
-  where the preorder $\future$ distributes to the underlying preorder, i.e.\
-  \[
-\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
-  \]
-  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
-  \[
-\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
-  \]
-\end{definition}
-\begin{definition}[Union preordered c.o.f.e.]
-  \label{def:sum-p-cofe}
-  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
-  \[
-  \left(X \cup Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
-  \]
-  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
-  \[
-    \text{for } z,z' \in X \cup Y,\quad  z' \future z \text{ iff } \left\{
-      \begin{array}{l}
-        z,z' \in X \wedge z' \future_x z \vee{} \\
-        z,z' \in Y \wedge z' \future_Y z
-      \end{array}
-    \right.
-  \]
-  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
-  \[
-    \text{for } z,z' \in X \cup Y,\quad  z \nequal z' \text{ iff } \left\{
-      \begin{array}{l}
-        z,z' \in X \wedge z \nequal_x z' \vee{} \\
-        z,z' \in Y \wedge z \nequal_Y z'
-      \end{array}
-    \right.
-  \]
-\end{definition}
+There are also natural constructions for preordered c.o.f.e.'s that rely on the underlying preorder.
+We will need the standard construction for later, product, and sum (defined in \appendixname~\ref{app:cofe}).
+
 Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
 
-In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get the world equation that we want.
-First, we define the c.o.f.e.\ that we need for the memory invariants.
-Note that we do not need a preorder on this.
+In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
+First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
 As a reminder, we sketched the memory invariants to as follows
@@
 \]
-We will need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
-The world will be given to us as the solution to the recursive equation.
-Generally, we can lift a set of relations to a c.o.f.e.\ by adding an index to it that is downwards closed.
+To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
+We do not need a construction for the world as it will be given as the solution to the recursive equation.
+Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
 This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
@@
   \]
+  to get a preordered c.o.f.e.\, we use subset as the preorder.
 \end{definition}
@@
 \label{def:monnefun-cofe}
-  Given a c.o.f.e.
-$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
+  Given a preordered c.o.f.e.
+$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where
   \[
@@
 Next, we construct the seal invariant.
-As a reminder, we sketched the seal invariant to be
+As a reminder, we sketched the seal invariant as
 \[
@@
 \]
-In order to define this as a c.o.f.e., we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
-This means that we just need to lift the set of seals to a c.o.f.e.\ and define the c.o.f.e.\ partial functions.
+In order to define the c.o.f.e. for this, we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
+This leaves us with the task of lifting the set of seals to a c.o.f.e.\ and defining the c.o.f.e.\ partial functions.
 Generally, all countable sets can be lifted to a c.o.f.e.\ with a family of equivalences that has full precision for all indices except index 0.
@@
 
-The c.o.f.e. of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions
+The c.o.f.e.\ of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions.
 \begin{definition}[The c.o.f.e. of partial, non-expansive functions]
@@
 Next, we want to define the regions of the world.
-As mentioned, regions have the added structure of a future region relation, so we need to define them as a preordered c.o.f.e.
-The future region relation was presented in Section~\ref{subsubsec:ft-and-revocation}, so we just need to use it for our definitions.
-The future world relation for the shared region is trivial as it is just equality
-The shared regions we sketched are the easiest to define in terms of a c.o.f.e.\ as the future relation is just equality.
+As mentioned, our regions have the added structure of the future region relation, so we need to define them as a preordered c.o.f.e.
+The future region was presented in Section~\ref{subsubsec:ft-and-revocation}.
+On shared regions, the relation is equality which makes it straight forward to define:
 \begin{definition}[Shared region preordered c.o.f.e.\ ]
@@
   \[
-  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
-  \]
-  where the preorder $\future$ is equality, and the family of equivalences distributes to the underlying c.o.f.e.'s, i.e.
-given $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
+  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},=\right)
+  \]
+  where the family of equivalences distributes to the underlying c.o.f.e.'s, i.e. given shared regions $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
   \[
@@
 
-Next, we define the preordered c.o.f.e.\ for shared regions.
-The future region relation on shared regions is non-trivial, so the definition of the preordered c.o.f.e.\ is a bit more involved.
-\begin{definition}[Shared region preordered c.o.f.e.\ ]
-  \label{def:shared-reg-p-cofe}
+Next, we define the preordered c.o.f.e.\ for spatial regions.
+The future region relation on spatial regions is non-trivial, so the definition of the preordered c.o.f.e.\ is slightly more involved.
+\begin{definition}[Spatial region preordered c.o.f.e.\ ]
+  \label{def:spatial-reg-p-cofe}
   Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
@@
 \end{definition}
-Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:shared-reg-p-cofe} define the preoreded c.o.f.e.'s necessary for the regions in our worlds.
+Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:spatial-reg-p-cofe} define the preordered c.o.f.e.'s necessary for the regions in our worlds.
 
@@
 \end{definition}
-The three sub-worlds we need can be obtained from Definition~\ref{def:sub-world-p-cofe}.
-First, we define the free stack sub-world simply by using the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) with the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
+The free stack sub-world preordered c.o.f.e.\ is constructed from the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) and the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
 The result is the sub-world for the free stack
@@
 Next, we define the call stack sub-world.
-To this end, we use Definition~\ref{def:shared-reg-p-cofe} to get the shared region.
-We need to combine this with an address.
-We achieve this by first lifting the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combining the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with Definition~\ref{def:product-p-cofe}.
+To this end, we use Definition~\ref{def:spatial-reg-p-cofe} to get the spatial region which we need to combine this with an address.
+We lift the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combine the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with the standard product construction.
 All in all, we get the a preordered c.o.f.e.\
@@
 Finally, we define the sub-world for the heap.
-To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the union preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
+To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) using the standard union preordered c.o.f.e.\ construction.
 The result is a preordered c.o.f.e.\
@@
 
-With the three sub-worlds, we can combine them to get our worlds using the preordered c.o.f.e.\ construction for products (Definition~\ref{def:product-p-cofe}) which gives us the preordered c.o.f.e.
+With the three sub-worlds defined, we combine them with the standard product construction for preordered c.o.f.e.'s to get our worlds.
 \[
@@
 This is the world, we would like to work with, but we still don't have a solution for the recursive domain equation.
-In fact, when we defined the regions, we just specified where the solution to the recursive domain equation should go.
-So what we previously sketched to be the recursive occurrence of the world is now a variable in the above.
+In fact, when we defined the regions, we just left world-index in the memory and seal invariants as an unspecified variable, which means that the above world has them as a variable.
 In order to find a solution, we still need to guard the recursive occurrence.
@@
 Now the recursive occurrence is guarded which means that there is a solution to the recursive domain equation.
-
 \begin{theorem}

### 2019-08-08 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add Iris remark in solving recursive domain equation section

@@
 This section does not provide further intuition about the worlds and can be skipped.
+
+Recursive domain equations are so common and the solution so standard that frameworks implement them lifting the technical burden of constructing the world by hand.
+The technique we use to solve our recursive domain equation is essentially the same used internally in the program logic framework Iris~\citep{iris1,iris2,iris3} to construct worlds.
+We believe the worlds presented here could be developed in Iris.
+Had we done the development in Iris Coq, then we would have had the world tool support of Iris and the rigorousness of Coq which, in hindsight, would have been very beneficial.
 

### 2019-08-08 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add a bit of intuition to the code region description

@@
 
-In our results, we assume that we execute well-formed components which allows us to make certain assumptions about the code we execute.
-We express these assumption in what we call a \emph{code region}.
-The invariant of the code region is defined as
+In our result, we assume well-formed components which puts certain syntactic constraints on the components.
+We also have the semantic assumption that trusted components are reasonable.
+Both assumptions need to be captured in the logical relation in order for us to rely on them.
+To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components..
+The memory invariant of the code region is defined as
 \begin{multline*}
@@
 \end{multline*}
-The code region is a lot more restrictive than the standard region.
-It basically only allows one memory segment, namely $\code$ padded with zeroes to make sure that two capabilities cannot be spliced to cause unintended control-flow.
-We will use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
+The code region is more restrictive than the standard region.
+It only allows one memory segment, namely $\code$ padded with zeroes that make sure that two capabilities cannot be spliced to cause unintended control-flow.
+We use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
 The assumptions we can make on the code depends on whether it is part of a trusted or untrusted component.

### 2019-08-08 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Update condition naming

@@
 
-We capture the essence of what it means for a capability with read permission to be safe in the condition $\readCond{}$.
+We capture the essence of what it means for a capability with read permission to be safe in the condition $\readCond[]{}$.
 % each of the regions upper bounded by std region
-The main purpose of $\readCond{}$ is to make sure that only safe words can be read from the memory governed by a read capability..
+The main purpose of $\readCond[]{}$ is to make sure that only safe words can be read from the memory governed by a read capability..
 This is done by putting an upper bound on what requirements an invariant can impose on the memory segments governed by the capability.
@@
 % splicing and what part of the range each region governs
-The $\readCond{}$ is compatible with all the operations that can be performed on capabilities.
-This means that if two capabilities, for which $\readCond{}$ holds, are spliced together, we can establish that $\readCond{}$ holds for the resulting capability.
+The $\readCond[]{}$ is compatible with all the operations that can be performed on capabilities.
+This means that if two capabilities, for which $\readCond[]{}$ holds, are spliced together, we can establish that $\readCond[]{}$ holds for the resulting capability.
 To support this, we require the presence of a set of regions $S$ that governs the addresses the capability has authority over rather than just a single region.
-If we need to establish the $\readCond{}$ after a splice, we can simply use the union of the regions that witnessed the $\readCond{}$ of the two individual capabilities.
-We also need to support splitting which is no problem for normal capabilities as the same $\pure$ region can be used to establish the $\readCond{}$ for multiple normal capabilities.
-On the other hand, a $\spatialo$ region can only be used to establish the $\readCond{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
-None the less, we need to support arbitrary splitting of linear capabilities, which means that $\readCond{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond{}$.
-This is why, $\readCond{}$ requires all regions to only govern one address when the capability is linear.
+If we need to establish the $\readCond[]{}$ after a splice, we can simply use the union of the regions that witnessed the $\readCond[]{}$ of the two individual capabilities.
+We also need to support splitting which is no problem for normal capabilities as the same $\pure$ region can be used to establish the $\readCond[]{}$ for multiple normal capabilities.
+On the other hand, a $\spatialo$ region can only be used to establish the $\readCond[]{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
+None the less, we need to support arbitrary splitting of linear capabilities, which means that $\readCond[]{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond[]{}$.
+This is why, $\readCond[]{}$ requires all regions to only govern one address when the capability is linear.
 This means that after a split, the authority of the regions for the bottom half of the split can go to one capability and the remaining regions can go to the top half.
@@
 
-Similarly to $\readCond{}$, we define a condition that captures the essence of what it means for a capability with writer permission to be safe.
-We call this condition $\writeCond{}$.
+Similarly to $\readCond[]{}$, we define a condition that captures the essence of what it means for a capability with writer permission to be safe.
+We call this condition $\writeCond[]{}$.
 A capability with write permission can be used to write to memory.
@@
 Say, you manage to write something that can break memory invariants, then it would not be possible to read it back again as write permission, generally speaking, does not entail read permission.
-If the capability had read permission, then $\readCond{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants.
+If the capability had read permission, then $\readCond[]{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants.
 After all, if such a capability was obtained, memory invariants could be broken.
-  However, the $\writeCond{}$ tries to capture the essence of safety and in principle it is safe to write an unsafe capability that cannot be read back.}.
+  However, the $\writeCond[]{}$ tries to capture the essence of safety and in principle it is safe to write an unsafe capability that cannot be read back.}.
 It should always be possible to write safe values, so we impose this as a lower bound.
@@
 \end{definition}
-The definition of $\writeCond{}$ is very similar to $\readCond{}$.
+The definition of $\writeCond[]{}$ is very similar to $\readCond[]{}$.
 Support for split and splice is done in the same way, and the bound is defined in terms of the standard region.
 
-The $\readCond{}$ and $\writeCond{}$ specifically uses the heap sub-world which means that it can only be used for heap capabilities.
+The $\readCond[]{}$ and $\writeCond[]{}$ specifically uses the heap sub-world which means that it can only be used for heap capabilities.
 This means that we cannot use it for stack capabilities.
-To take care of stack capabilities, we define two more conditions a $\stackReadCond{}$ and $\stackWriteCond{}$.
-The two new condition are essentially the same as the $\readCond{}$ and $\writeCond{}$ except that they use the free stack sub-world and assume that the capability is linear as all stack capabilities are linear.
+To take care of stack capabilities, we define two more conditions a $\stackReadCond[]{}$ and $\stackWriteCond[]{}$.
+The two new condition are essentially the same as the $\readCond[]{}$ and $\writeCond[]{}$ except that they use the free stack sub-world and assume that the capability is linear as all stack capabilities are linear.
 Note that we do not have any condition that talks about the stack-frames sub world because we should never have a capability that allows us to directly read from or write to that part of memory.
@@
 The final permission, we define conditions for is the execute permission.
-We define two conditions $\execCond{}$ and $\xReadCond{}$.
-The $\execCond{}$ captures what operations an execute-capability can be used for, i.e.\ execution.
-The $\xReadCond{}$ captures some additional read assumptions we can make on a capability when we know the capability is executable.
-
-The $\execCond{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
+We define two conditions $\execCond[]{}$ and $\xReadCond[]{}$.
+The $\execCond[]{}$ captures what operations an execute-capability can be used for, i.e.\ execution.
+The $\xReadCond[]{}$ captures some additional read assumptions we can make on a capability when we know the capability is executable.
+
+The $\execCond[]{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
 We later define the $\lrename$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
@@
 For this reason, the condition requires that any executable capability with a derived range of authority and a current address in that range is safe to use for execution.
-The $\execCond{}$ is quantified over all future worlds of the $\purePart{}$ of $W$.
+The $\execCond[]{}$ is quantified over all future worlds of the $\purePart{}$ of $W$.
 We do not know when the executable capability will be used, so it should be safe even in the future when the memory has changed.
@@
 \end{align*}
-The $\readCond{}$ condition by itself allows many different regions and thus potentially many different memory segments.
+The $\readCond[]{}$ condition by itself allows many different regions and thus potentially many different memory segments.
 However, when we have a read capability with execute permission, we know that the capability must point to a piece of code memory.
-For this reason, we define the $\xReadCond{}$ to capture the additional assumptions that we can make when a capability is executable.
+For this reason, we define the $\xReadCond[]{}$ to capture the additional assumptions that we can make when a capability is executable.
 \[
@@
 \]
-The $\xReadCond{}$ requires that the memory segment an executable capability has authority over is governed by a code region.
-Note that we do not define $\execCond{}$ and $\xReadCond{}$ for the stack because the stack is not executable.
+The $\xReadCond[]{}$ requires that the memory segment an executable capability has authority over is governed by a code region.
+Note that we do not define $\execCond[]{}$ and $\xReadCond[]{}$ for the stack because the stack is not executable.
 
@@
 This condition and conditions for the other permissions are captured by the permission based conditions, so we use them to express the necessary conditions.
-That is, if a capability has read permission, then $\readCond{}$ must be satisfied, if it has write permission, then $\writeCond{}$ must be satisfied, and if it has execute permission, then $\execCond{}$ and $\xReadCond{}$ must be satisfied.
+That is, if a capability has read permission, then $\readCond[]{}$ must be satisfied, if it has write permission, then $\writeCond[]{}$ must be satisfied, and if it has execute permission, then $\execCond[]{}$ and $\xReadCond[]{}$ must be satisfied.
 If the capability has execute permission, then it must also be a normal capability.
@@
 Two capabilities are related as trusted code pointers if they are normal, has a permission derivable from read-execute, and has a range of authority within the trusted address space, $\ta$.
-Further, we need to know that the capabilities actually point to a piece of code which is why we require the $\xReadCond{}$ to be satisfied.
+Further, we need to know that the capabilities actually point to a piece of code which is why we require the $\xReadCond[]{}$ to be satisfied.
 This makes sure that the region that governs the memory the capability points to is a code region.

### 2019-08-08 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Write some more about trusted value relation

@@
 A value relation based on what it is safe for an adversary to have should prohibit return seals, so such a relation cannot be used to reason about trusted programs.
-For this reason, we define two value relations a trusted $\lrvname_{\trusted}$ and an untrusted $\lrvname$.
+For this reason, we define two value relations a trusted $\lrvname_{\trusted}$ and an untrusted $\lrvname_{\untrusted}$.
 Anything safe for unstrusted programs is also safe to give to a trusted program, so the trusted value relation is defined as a super set of the untrusted value relation.
@@
 The untrusted and trusted value relation both relates \srccm{} and \trgcm{} words.
-The untrusted value relation $\lrvname$ relates words that are safe to give to untrusted programs and $\lrvname_\trusted$ relates words that are safe to give to trusted programs.}
+The untrusted value relation $\lrvname_{\untrusted}$ relates words that are safe to give to untrusted programs and $\lrvname_\trusted$ relates words that are safe to give to trusted programs.}
 \label{fig:value-relation}
@@
 \label{par:untrusted-val-rel}
-The untrusted value relation $\lrvname$ relates all the words that untrusted components can safely posses.
+The untrusted value relation $\lrvname_\untrusted$ relates all the words that untrusted components can safely posses.
 That is words that cannot be used to break any memory invariants.
@@
 
+The words in $\lrvname_\trusted$ but not in the $\lrvname_\untrusted$ have the potential to break the system invariants \stktokens{} rely on.
+We can only let trusted components have words from $\lrvname_\trusted$ because the trusted component promises to not break the invariant by behaving reasonably.
+This promise is expressed formally in $\lrvname_\trusted$ by requiring the presence of a code region in both cases specific to $\lrvname_\trusted$.
+As explained previously, the code region essentially captures the requirements put on components by well-formedness and the reasonability condition which constitutes the promise to use seals and trusted code pointers in a way that does not break invariants.
+
 % Trusted seals
@@
 Note that even though the capability has read permission, we do not require the read condition to hold.
-The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrvname$, so the read condition would not hold.
+The code memory contains trusted closure seals and return seals that we cannot let untrusted code have and the read condition requires everything to be in $\lrvname_\untrusted$, so the read condition would not hold.
 However, trusted code can have access to such seals because we expect the trusted code to treat the seals reasonably.
@@
 \end{figure}
+Like the untrusted value relation, the trusted value relation is monotone.
+Intuitively, the two relations are monotone for the same reason; words are potentially used at any point in time.
+If words are safe now (in the current world), then they should also be safe to use later (in any possible future world).
+\begin{lemma}[Trusted value relation monotonicity]
+  \label{lem:monotonicity-trusted}
+  For all integers $n$, words $\src{w_1}$ and $w_2$, and worlds $W' \future W$, if $\npair{(\src{w_1},w_2)} \in \lrvtrusted(W)$, then $\npair{(\src{w_1},w_2)} \in \lrvtrusted(W')$.
+\end{lemma}
 

### 2019-08-08 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add "non-linear is pure" lemma

@@
 \end{lemma}
+Another, perhaps unsurprising property, of the trusted value relation it that non-linear words do not depend on the spatial region that may be in the world.
+This is unsurprising as normal capabilities do not necessarily reference memory uniquely.
+\begin{lemma}[Non-linear words are independent of spatial regions]
+  \label{lem:non-linear-pure}
+ If $\npair{(w_1,w_2)} \in \lrvtrusted(W)$ and either $\nonLinear{w_1}$ or $\nonLinear{w_2}$, then
+ \[
+   \npair{(w_1,w_2)} \in \lrvtrusted(\purePart{W})
+ \].
+\end{lemma}
+There is similar lemma to Lemma~\ref{lem:non-linear-pure} for the untrusted value relation.
 

### 2019-08-09 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Refactor untrusted val rel sealed case

@@
 
+The $\execCond[]{}$ handles normal jumps, but it does not cover the case of $\txjmp{}{}$.
+Executable capabilities can be used on their own whereas sealed capabilities must be jumped to in pairs.
+However, we do not need to consider arbitrary pairs: given a sealed capability we only have to consider the capabilities permitted by the relevant seal invariant.
+Just like the $\lrename$ relation captures what it means for a word to be safe as a program counter, we later define $\lrexjname$ that define what it means for a code and data capability pair to be safe together as program counter and data capability, respectively.
+A sealed capability is safe when it can be paired with any sealed capability from the seal invariant such that the pair is in the $\lrexjname$ relation.
+Just like safe executable capabilities, a sealed capability may be stored, so it should also be safe to use in future worlds.
+The condition for sealed capabilities is defined by $\sealedCond[]{}$.
+\begin{multline*}
+\sealedCond{W,H_\sigma} =\\
+\left\{\npair{(\sigma, \src{\vsc_S},\vsc_T)}\middle|
+  \begin{array}{l}
+  \forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp\\
+    \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
+  \end{array}
+\right\}
+\end{multline*}
+
 \paragraph{The untrusted value relation}
@@
 The $\xjmp{}{}$ takes a pair of sealed capabilities, so the authority of a sealed capability depends on what other sealed capabilities it might be used with.
-The seal invariant defines what may be sealed with a given seal and thus which sealables we should consider.
-As discussed, closure seals and return seals must be used in specific ways, so we have defined a seal invariant that dictates what may be sealed with a given seal.
-In order for a pair of \srccm{} and \trgcm{} seals to be in the untrusted value relation, they must be sealed with the same seal $\sigma$, and they must be related by the seal invariant $\sigma$.
+The seal invariant specifies the capabilities that may be sealed with a given seal and thus the capabilities that may be used together as a sealed pair.
+As discussed, closure and return seals must be used in specific ways which is captured in the code region seal invariant.
+In order for a pair of \srccm{} and \trgcm{} sealed capabilities to be in the untrusted value relation, they must be sealed with the same seal $\sigma$ and related in the appropriate seal invariant.
 % Sealed capability only useful with pairing. Sealed caps must be permitted.
-Further, if we take any sealed capabilities related under $\sigma$ in the seal invariant, then they should produce a safe result.
-We are yet to define what it means for an $\txjmp{}{}$ to produce a safe result, but we will do so later with the $\lrename_\mathrm{xjmp}$-relation which basically says that if an x-jump happened with the capabilities in question, then the resulting configuration is in the observation relation, i.e.\ it respects the memory invariants.
-% Linear and non-linear case. Affects the future world we consider.
-It should be safe to use the sealed capability both now and in the future.
-This means that it should be safe in the current world $W$, but also in future worlds $W' \future W$.
-However, this only goes for non-linear words.
-For linear capabilities, it should only be in future worlds of the shared part of $W$.
-% TODO write why we have the two cases. (it allows us to have the "other" capability by the one we consider - and it feels like the only reason)
+Further, they should satisfy the $\sealedCond[]{}$ which means that they can safely be paired up with any other pair of capabilities from the seal invariant and used safely for execution.
 
@@
       (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
-      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
+      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \\
+      \quad\pwheap(r) = (\pure,\_,H_\sigma) \tand \\
       \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
       \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
-      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
-      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
-      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
-
+      \quad \isLinear{\src{\vsc_S}} \\
+      \qquad \Rightarrow \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in \sealedCond{W,H_\sigma} \wedge\\
+      \quad \nonLinear{\src{\vsc_S}} \\
+      \qquad \Rightarrow  \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in \sealedCond{\mathit{purePart}(W),H_\sigma}\\
       % This case was never used (?) so it is not necessary.

### 2019-08-09 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Reformulate paragraph

@@
 Both assumptions need to be captured in the logical relation in order for us to rely on them.
-To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components..
+To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components.
 The memory invariant of the code region is defined as
@@
 % each of the regions upper bounded by std region
-The main purpose of $\readCond[]{}$ is to make sure that only safe words can be read from the memory governed by a read capability..
+The main purpose of $\readCond[]{}$ is to make sure that only safe words can be read from the memory governed by a read capability.
 This is done by putting an upper bound on what requirements an invariant can impose on the memory segments governed by the capability.
@@
 
-% TODO the following paragraph does not read well. Rewrite.
-The upper bound on requirements that can be imposed on the memory the read capability has access to is specified on all of the regions that govern the memory.
-The upper-bound is specified in terms of the standard region $\stdreg{R(r),\gc}{\pur}$.
-That is, the memory segments permitted by the region should be an $n$-subset (a subset that only takes into account elements with an index up to, but not including, $n$).
-This means that everything the region permits must at the very least be safe.
+A safe read capability only gives authority to read safe words.
+The invariant on the memory a read capability gives access to may be even more restrictive than just requiring safe words.
+For instance, the invariant may require a flag to stay unchanged.
+We express the fact that a region may be more restrictive by making the standard region $\stdreg{R(r),\gc}{\pur}$, which permits all memory segments with safe words, the upper bound of what a region may require when it governs a memory segment that can be accessed through a safe read capability.
 

### 2019-08-09 popl-2019-journal-v/paper.tex
[POPLjv] Write contribution outline for the introduction

@@
 \begin{jversion}
-  % TODO add outline of new work
   This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens_2019}.
-  Compared to the earlier text, this version adds X and Y and Z.
+  Compared to the earlier text, this version adds and explains many of the details that were left out in the conference version due to space restrictions.
+  The added details include a better motivation of sealing (\sectionname~\ref{sec:purpose-sealing}), well-formedness components (\sectionname~\ref{sec:well-form-reas}), and reasonable components (\sectionname~\ref{sec:reasonable-components}).
+  The section on proving full-abstraction (\sectionname~\ref{sec:fa-proof}) has been rewritten with details and explanations about the formal method used for the full-abstraction proof.
+  This includes a detailed description and motivation of the world (\sectionname~\ref{subsec:worlds}) and logical relation (\sectionname~\ref{subsec:logical-relation}).
   This paper is accompanied by a technical report~\citep{technical_report_popl} with the elided details and proof.
 \end{jversion}
+% Notes on new contributions (based on the journal version environment):
+% More details about the definition of the cap. machine (decoding/encoding functions)
+% Well-formedness definitions
+% Reasonability definition
+% Expanded discussion section
+% More detailed description of worlds and
+%   the logical relation used
 
@@
 \subsection{Reasonable Components}
+\label{sec:reasonable-components}
 % Reasonability

### 2019-08-09 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Fix bibtex

@@
 Recursive domain equations are so common and the solution so standard that frameworks implement them lifting the technical burden of constructing the world by hand.
-The technique we use to solve our recursive domain equation is essentially the same used internally in the program logic framework Iris~\citep{iris1,iris2,iris3} to construct worlds.
+The technique we use to solve our recursive domain equation is essentially the same used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
 We believe the worlds presented here could be developed in Iris.

### 2019-08-11 popl-2019-journal-v/paper.tex
Add important point to discussion of Tsampas et al.
typo

@@
 % KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive.
+However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearingt the entire unused part of the stack, an operation that may be prohibitively expensive.
 
@@
 This means that their approach is more like C than what \stktokens{} has to offer.
+As explained in Section~\ref{sec:introduction}, these approaches have the downside that they require stack clearing (including unused parts) on boundary crossings.
 

### 2019-08-11 popl-2019-journal-v/paper.tex
Do a pass over discussion of robust hyperproperty preservation paper.

@@
 
-There are other notion of secure compilation than full-abstraction~\citep{abadi_protection_1998}.
-\citet{abate_2019} presents an overview of trace-based secure compilation properties.
-They show that full-abstraction is a weak notion of secure compilation relatively to other notions, e.g., Robust Relational Hyperproperty Preservation (RrHP) their strongest notion of secure compilation.
-However, other notions of secure compilation that promise WBCF and LSE must also enforce it somehow.
-\stktokens{} is an enforcement scheme that likely would work for stronger notions of secure compilation.
-In fact, we believe that our result could be lifted from full-abstraction to RrHP reusing the back-translation from the full-abstraction proof.
+There are other notions of secure compilation than full-abstraction~\citep{abadi_protection_1998}.
+\citet{abate_2019} present an overview of trace-based secure compilation properties.
+Full abstraction is only one, relatively weak, property in their hierarchy.
+It would be interesting to investigate if our compiler, i.e.\ the embedding function from \srccm{} into \trgcm{}, also satisfies some of their other properties.
+While our current result implies that we can prove contextual equivalences in \trgcm{} components using \stktokens{}, by proving them instead in the more well-behaved \srccm{} semantics, such stronger properties would imply that we can also prove robust preservation of other (hyper-)properties in a similar manner.
+We expect that, in addition to full abstraction, our embedding also satisfies Robust Relational Hyperproperty Preservation (RrHP, the strongest property in the hierarchy of \citeauthor{abate_2019}) and that a large part of our current proof (the back-translation and the logical relation) could be reused to establish this.
+However, to do this, we would first need to extend our semantics with some form of traces and we have not investigated how best to do this.
 

### 2019-08-11 popl-2019-journal-v/paper.tex
Justify our expectation

@@
 While our current result implies that we can prove contextual equivalences in \trgcm{} components using \stktokens{}, by proving them instead in the more well-behaved \srccm{} semantics, such stronger properties would imply that we can also prove robust preservation of other (hyper-)properties in a similar manner.
-We expect that, in addition to full abstraction, our embedding also satisfies Robust Relational Hyperproperty Preservation (RrHP, the strongest property in the hierarchy of \citeauthor{abate_2019}) and that a large part of our current proof (the back-translation and the logical relation) could be reused to establish this.
+As our back-translation works for arbitrary programs, we expect that, in addition to full abstraction, our embedding also satisfies Robust Relational Hyperproperty Preservation (RrHP, the strongest property in the hierarchy of \citeauthor{abate_2019}) and that a large part of our current proof (the back-translation and the logical relation) could be reused to establish this.
 However, to do this, we would first need to extend our semantics with some form of traces and we have not investigated how best to do this.

### 2019-08-11 popl-2019-journal-v/proving-full-abstraction.tex
Minor fixes in discussion of LR

@@
 \end{lemma}
-Another, perhaps unsurprising property, of the trusted value relation it that non-linear words do not depend on the spatial region that may be in the world.
+Another, perhaps unsurprising property, of the trusted value relation it that non-linear words do not depend on the spatial regions that may be in the world.
 This is unsurprising as normal capabilities do not necessarily reference memory uniquely.
@@
  \[
-   \npair{(w_1,w_2)} \in \lrvtrusted(\purePart{W})
- \].
+   \npair{(w_1,w_2)} \in \lrvtrusted(\purePart{W})\text{.\qedhere}
+ \]
 \end{lemma}
-There is similar lemma to Lemma~\ref{lem:non-linear-pure} for the untrusted value relation.
+This is similar to Lemma~\ref{lem:non-linear-pure} for the untrusted value relation.
 

### 2019-08-12 toplas/paper.tex
[TOPLAS] Add reviews with todo and first round of fixes

@@
 They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that have recently received renewed interest.
-in particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
+In particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
 Capability machines tag every word (in the register file and in memory) to enforce a strict separation between numbers and capabilities (a kind of pointers that carry authority).
@@
 The future world relations model how memory may evolve over time.
-We have a \emph{public} and a \emph{private} future world relation that respectively what changes any capability can rely on and what changes only temporary regions can rely on.
-As temporary capabilities fall within the category of all capabilities, the public future relation is subsumed in the private future relation.
+We have a \emph{public} and a \emph{private} future world relation that, respectively, model the memory changes any capability can rely on and the memory changes only local capabilities can rely on.
+As local capabilities fall within the category of all capabilities, the public future relation is subsumed in the private future relation.
 
@@
 The logical relation defines semantically when values, program counters, and configurations are capability safe.
-The logical relation is defined found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and we provide some explanations in the following paragraphs.
+The logical relation is defined in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and we provide some explanations in the following paragraphs.
 For space reasons, we omit some definitions and explain them only verbally, but precise definitions can be found in the appendix.
@@
 In this case, it is a \tpwl{} $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
-The \tpwl{} only accept memories with address range $[\start,\addrend]$.
+The \tpwl{} only accepts memories with address range $[\start,\addrend]$.
 More importantly, the \tpwl{}'s state interpretation function only accepts memory segments that only contain safe words.
@@
 This is expressed in $\readCond{}$ with $\var{localityReg}(\gl,W)$ which projects out all the regions the capability can depend on based on its locality.
-Specifically, if the capability is local, then all active regions (non-revoked) are project.
+Specifically, if the capability is local, then all active regions (non-revoked) are projected.
 On the other hand, if the capability is global, then only the permanent regions are projected.
@@
   In our definition of worlds, nothing prevents a world from having regions that are overlapping.
-  This might seem like an issue the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
+  This may seem like an issue with the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
   In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness condition on the world as it prevents the regions from overlapping.
@@
   However, formulating these requirements is harder than one might expect.
-  The problem is that a realistic implementation of malloc needs to rely on internal state that changes after every invocation and relies on invariants on that state.
+  The problem is that a realistic implementation of malloc needs to rely on internal state (the free memory is part of malloc's internal state) that changes after every invocation and relies on invariants on that state.
   We can only expect that malloc behaves according to its specification if its internal state satisfies the current state of the invariants in an executing system.
@@
 macro reduces the range of authority of the stack pointer to the unused part of
-the stack and clears it(as discussed in
-Section~\ref{sec:stack-and-return-pointer}). Finally, the \texttt{scall} macro clears
+the stack and clears it (as discussed in Section~\ref{sec:stack-and-return-pointer}). Finally, the \texttt{scall} macro clears
 non-argument registers and jumps to $r$. Upon return after the restoration code
@@
 It is especially annoying when the same line of reasoning is applied in multiple places.
-To mitigate this, we have define a number of lemmas that capture common reasoning patterns in these proofs.
+To mitigate this, we have defined a number of lemmas that capture common reasoning patterns in these proofs.
 In this section, we present the most central lemmas used to reason about assembly programs.
@@
     \item $W' =
-      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
+      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$\footnote{We use the update-notation without a region name to indicate that it should be a fresh region name.},
     \item $\memSat[n-1]{\ms''}{W'}$
@@
 As explained previously, we use the FTLR (Theorem~\ref{thm:ftlr}) to argue about unknown code; this is no exception.
-That is, when we argue that we assumption~\ref{item:scall:hyp-callee} in the \texttt{scall} lemma satisfied, we use the FTLR along with the assumptions we have (e.g.\ linking table, malloc capability etc.).
+That is, when we argue that assumption~\ref{item:scall:hyp-callee} in the \texttt{scall} lemma is satisfied, we use the FTLR along with the assumptions we have (e.g.\ linking table, malloc capability etc.).
 In Hyp-Callee, we assume the memory is safe, so it suffices to show that the register-file contains safe values, which amounts to showing that the arguments in the call are safe.
@@
 % capability machine.
-In principle, one could reason solely in the semantic model
+% In principle, one could reason solely in the semantic model
 % without the use of these lemmas, but that would lead to a lot of repetition.
@@
 Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
-If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
+If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $0$, causing the assertion to fail.
 Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
@@
   Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{cf.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
-  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.
+  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked.
   In $W_2$, two new regions govern the stack.
@@
 \emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
-Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
+Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, \itoplas{creates a view with all capabilities that may be available at some point in time.
+By removing the time aspect, we may have to consider capabilities that are actually not available at a given point in time.}
   Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
@@
 The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
-They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
+They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behave as though there was a built-in stack.
 Their full abstraction proof uses a logical relation, but it differs in a number of ways from the one we present.
@@
   By making these simplifications, we avoid tedious details but end up with a less realistic machine.
-  However, the intend of this work is to gain ground in the formal work necessary to prove properties about low-level assembly programs on a capability machine not to apply it to a full fledged capability machine.
+  However, the intent of this work is to gain ground in the formal work necessary to prove properties about low-level assembly programs on a capability machine not to apply it to a full fledged capability machine.
 
   We do not believe that any of our simplifications are beyond reason and expanding the result seems plausible.
-  If we added unbounded integers, we would have to change the operational
+  If we added bounded integers, we would have to change the operational
   semantics to take over and under flows into account. This could be achieved
@@
 the registers are still confined to that range of memory. The property talks
-about a the specific implementation of a higher-order value rather than its behavior,
+about the specific implementation of a higher-order value rather than its behavior,
 like the invariants that are required/preserved when we use it.

### 2019-08-12 toplas/appendix.tex
[TOPLAS] Push appendix tex file

--- 2019-03-06 toplas/paper.tex
@@
-\documentclass[format=acmsmall, review=true, screen=true]{acmart}
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-
-% Metadata Information
-\acmJournal{TWEB}
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2010}
-\acmMonth{3}
-\copyrightyear{2009}
-%\acmArticleSeq{9}
-% Copyright
-%\setcopyright{acmcopyright}
-\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{0000001.0000001}
-
-% Paper history
-%\received{February 2007}
-%\received[revised]{March 2009}
-%\received[accepted]{June 2009}
-
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\usepackage{cite}
-\usepackage{wrapfig}
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage{thmtools}
-\usepackage{stmaryrd}
-% \usepackage[numbers]{natbib}
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{array}
-\usepackage{arydshln}
-\usepackage{ifthen}
-\usepackage{ifpdf}
-\usepackage{verbatim}
-\usepackage{mathpartir}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{multicol}
-\lstset{
-  basicstyle=\footnotesize\ttfamily,
-  columns=fullflexible,
-  keepspaces=true,
-  mathescape,
-  escapeinside=||,
-  numbers=left,
-  numberblanklines=false,
-%  frame=single,
-%  framexleftmargin=12pt,
-  numbersep=5pt,
-  morecomment=[l]{//},
-  commentstyle=\color{blue},
-  morekeywords={push, pop, move, lea, restrict, subseg, geta, getb, mclear,
-    rclear, jmp, load, store, fetch, scall, assert, halt, malloc, call, crtcls,
-    reqglob, prepstk, plus, minus, jnz, gete}
-}
-
-\usepackage{rotating}
-\usepackage{afterpage}
-
-\usepackage{mathtools}
-\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
-\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
-
-\usepackage{enumitem}
-\newlist{enumproof}{enumerate}{10}
-\setlist[enumproof]{label*=\arabic*.}
-
-\usepackage{tikz}
-\usetikzlibrary{positioning,patterns,arrows,automata,decorations.markings}
-\usepackage{multirow,bigdelim}
-
-% \ifCLASSOPTIONcompsoc
-% \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-% \else
-% \usepackage[caption=false,font=footnotesize]{subfig}
-% \fi
-
-% \newtheorem{theorem}{Theorem}
-% \newtheorem{lemma}{Lemma}
-\input{preamble-illu}
-
-\renewcommand{\figurename}{Figure}
-\renewcommand{\sectionname}{Section}
-\newcommand{\forcenewline}{$\phantom{v}$\\}
-\newcommand{\judgment}[2]{\paragraph{#1}\hspace{\stretch{1}}\fbox{$#2$}}
-
-\newcommand{\update}[2]{[#1 \mapsto #2]}
-\newcommand{\sem}[1]{\left\llbracket #1 \right\rrbracket}
-% Math notation
-\newcommand{\restrictfun}[1]{|_{#1}}
-\newcommand{\parfun}{\rightharpoonup}
-\newcommand{\finparfun}{\xrightharpoonup{\textit{\tiny{fin}}}}
-\newcommand{\monnefun}{\xrightarrow{\textit{\tiny{mon, ne}}}}
-\newcommand{\monfun}{\xrightarrow{\textit{\tiny{mon}}}}
-\newcommand{\nefun}{\xrightarrow{\textit{\tiny{ne}}}}
-\newcommand{\fun}{\rightarrow}
-\newcommand{\defeq}{\stackrel{\textit{\tiny{def}}}{=}}
-\newcommand{\nequal}[1][n]{\stackrel{\tiny{#1}}{=}}
-\renewcommand{\nsim}[1][n]{\stackrel{\tiny{#1}}{\simeq}}
-
-\newcommand\subsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\subset$}\cr
-      \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
-\newcommand\supsetsim{\mathrel{\ooalign{\raise.2ex\hbox{$\supset$}\cr
-      \hidewidth\lower.8ex\hbox{\scalebox{0.9}{$\sim$}}\hidewidth\cr}}}
-\newcommand{\nsubsim}[1][n]{\stackrel{\tiny{#1}}{\subsetsim}}
-\newcommand{\nsupsim}[1][n]{\stackrel{\tiny{#1}}{\supsetsim}}
-
-\newcommand{\nsubeq}[1][n]{\stackrel{\tiny{#1}}{\subseteq}}
-\newcommand{\nsupeq}[1][n]{\stackrel{\tiny{#1}}{\supseteq}}
-
-\newcommand{\union}{\mathbin{\cup}}
-\DeclareMathOperator{\dom}{dom}
-\newcommand{\blater}{\mathop{\blacktriangleright}}
-\newcommand{\id}{\var{id}}
-% \newcommand{\undefined}{\mathit{undefined}}
-
-\newcommand{\powerset}[1]{\mathcal{P}(#1)}
-
-\newcommand{\false}{\mathit{false}}
-\newcommand{\true}{\mathit{true}}
-
-% cofes
-\newcommand{\cofe}{c.o.f.e.}
-\newcommand{\cofes}{\cofe{}'s}
-\newcommand{\CatC}{\mathbb{C}}
-\newcommand{\CatP}{\mathbb{P}}
-
-% Comments
-\newcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
-\newcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
-\newcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
-
-% Uncomment me to disable comments
-\renewcommand\lau[1]{}
-\renewcommand\dominique[1]{}
-\renewcommand\lars[1]{}
-
-% Variables
-\newcommand{\var}[1]{\mathit{#1}}
-\newcommand{\hs}{\var{ms}}
-\newcommand{\ms}{\hs}
-\newcommand{\hv}{\var{r}}
-\newcommand{\rv}{\var{r}}
-\newcommand{\lv}{\var{r}}
-\newcommand{\gl}{\var{g}}
-\newcommand{\pc}{\mathit{pc}}
-\newcommand{\pcreg}{\mathrm{pc}}
-\newcommand{\addr}{\var{a}}
-\newcommand{\offset}{\var{offset}}
-\newcommand{\word}{\var{w}}
-\newcommand{\start}{\var{b}}
-\newcommand{\addrend}{\var{e}}
-\newcommand{\pwlv}{\var{pwl}}
-\newcommand{\mem}{\var{mem}}
-\newcommand{\reg}{\var{reg}}
-\newcommand{\heapseg}{\var{ms}}
-\newcommand{\heap}{\var{mem}}
-\newcommand{\mode}{\var{mode}}
-\newcommand{\perm}{\var{perm}}
-\newcommand{\inftyend}{-42}
-\newcommand{\permp}{\var{permPair}}
-\newcommand{\roll}{\var{roll}}
-\newcommand{\instr}{\var{instr}}
-\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}
-\newcommand{\adv}{\var{adv}}
-\newcommand{\msframe}{ms_\var{frame}}
-\newcommand{\link}{\var{link}}
-\newcommand{\stk}{\var{stk}}
-\newcommand{\flag}{\var{flag}}
-\newcommand{\nwl}{\var{nwl}}
-\newcommand{\pwl}{\var{pwl}}
-\newcommand{\sta}{\var{sta}}
-\newcommand{\cnst}{\var{cnst}}
-\newcommand{\olf}{\var{offsetLinkFlag}}
-\newcommand{\prp}{\var{prp}}
-\newcommand{\env}{\var{env}}
-\newcommand{\cls}{\var{cls}}
-\newcommand{\unused}{\var{unused}}
-\newcommand{\act}{\var{act}}
-
-\newcommand{\hsframe}{\hs_\var{frame}}
-\newcommand{\size}{\var{size}}
-\newcommand{\rio}{r_{io}}
-\newcommand{\advb}{\var{adv_{base}}}
-\newcommand{\adve}{\var{adv_{end}}}
-\newcommand{\initb}{\var{init}_{base}}
-\newcommand{\inite}{\var{init}_{end}}
-\newcommand{\mrlen}{5cm}
-\newcommand{\retm}{\var{ret}_{\malloc}}
-\newcommand{\reta}{\var{ret}_{\adv}}
-\newcommand{\base}{\var{base}}
-\newcommand{\eend}{\var{end}}
-
-% Memory projections
-\newcommand{\plainproj}[1]{\mathrm{#1}}
-\newcommand{\memheap}[1][\Phi]{#1.\plainproj{mem}}
-\newcommand{\memreg}[1][\Phi]{#1.\plainproj{reg}}
-
-\newcommand{\updateHeap}[3][\Phi]{#1\update{\plainproj{mem}.#2}{#3}}
-\newcommand{\updateReg}[3][\Phi]{#1\update{\plainproj{reg}.#2}{#3}}
-
-% Configuration end states
-\newcommand{\failed}{\mathit{failed}}
-\newcommand{\halted}{\mathit{halted}}
-
-% Functions
-\newcommand{\plainfun}[2]{
-  \ifthenelse{\equal{#2}{}}
-  {\mathit{#1}}
-  {\mathit{#1}(#2)}
-}
-\newcommand{\decode}{\plainfun{decode}{}}
-\newcommand{\encode}{\plainfun{encode}{}}
-\newcommand{\encodePerm}{\mathit{encodePerm}}
-\newcommand{\encodePermPair}{\plainfun{encodePermPair}{}}
-\newcommand{\encodeLoc}{\mathit{encodeLoc}{}}
-\newcommand{\decodePermPair}{\plainfun{decodePermPair}}
-\newcommand{\decodePerm}[1]{\plainfun{decodePerm}{#1}}
-\newcommand{\decodeEnd}[1]{\plainfun{decodeEndAddr}{#1}}
-\newcommand{\decodeStart}[1]{\plainfun{decodeStartAddr}{#1}}
-\newcommand{\updatePcPerm}[1]{\plainfun{updPcPerm}{#1}}
-
-\newcommand{\executeAllowed}[1]{\plainfun{executeAllowed}{#1}}
-\newcommand{\nonZero}[1]{\plainfun{nonZero}{#1}}
-\newcommand{\readAllowed}[1]{\plainfun{readAllowed}{#1}}
-\newcommand{\writeAllowed}[1]{\plainfun{writeAllowed}{#1}}
-\newcommand{\withinBounds}[1]{\plainfun{withinBounds}{#1}}
-\newcommand{\stdUpdatePc}[1]{\plainfun{updPc}{#1}}
-
-\newcommand{\readCond}[1]{\plainfun{readCond}{#1}}
-\newcommand{\writeCond}[1]{\plainfun{writeCond}{#1}}
-\newcommand{\execCond}[1]{\plainfun{execCond}{#1}}
-\newcommand{\entryCond}[1]{\plainfun{enterCond}{#1}}
-
-\newcommand{\revokeTemp}[1]{\plainfun{revokeTemp}{#1}}
-\newcommand{\erase}[2]{\floor*{#1}_{\{#2\}}}
-\newcommand{\activeReg}[1]{\plainfun{active}{#1}}
-
-% World operations
-\newcommand{\future}{\mathbin{\sqsupseteq}}
-\newcommand{\pub}{\var{pub}}
-\newcommand{\priv}{\var{priv}}
-\newcommand{\futurewk}{\mathbin{\sqsupseteq}^{\var{pub}}}
-\newcommand{\futurestr}{\mathbin{\sqsupseteq}^{\var{priv}}}
-\newcommand{\heapSat}[3][\heap]{#1 :_{#2} #3}
-\newcommand{\memSat}[3][n]{\heapSat[#2]{#1}{#3}}
-\newcommand{\memSatPar}[4][n]{\heapSat[#2]{#1 , #4}{#3}}
-
-\newcommand{\monwknefun}{\xrightarrow[\text{\tiny{$\futurewk$}}]{\textit{\tiny{mon, ne}}}}
-\newcommand{\monstrnefun}{\xrightarrow[\text{\tiny{$\futurestr$}}]{\textit{\tiny{mon, ne}}}}
-
-% Assembly labels
-\newcommand{\codelabel}[1]{\mathit{#1}}
-\newcommand{\init}{\codelabel{init}}
-\newcommand{\malloc}{\codelabel{malloc}}
-\newcommand{\counter}{\codelabel{counter}}
-\newcommand{\iocap}{\codelabel{iocap}}
-
-% Type(s)
-\newcommand{\type}[1]{\mathrm{#1}}
-\newcommand{\asmType}{\plaindom{AsmType}}
-
-% Domains
-\newcommand{\plaindom}[1]{\mathrm{#1}}
-\newcommand{\Caps}{\plaindom{Cap}}
-\newcommand{\Words}{\plaindom{Word}}
-\newcommand{\Addrs}{\plaindom{Addr}}
-\newcommand{\ExecConfs}{\plaindom{ExecConf}}
-\newcommand{\RegName}{\plaindom{RegName}}
-\newcommand{\Regs}{\plaindom{Reg}}
-\newcommand{\Heaps}{\plaindom{Mem}}
-\newcommand{\Mems}{\Heaps}
-\newcommand{\HeapSegments}{\plaindom{MemSeg}}
-\newcommand{\MemSegments}{\HeapSegments}
-\newcommand{\Confs}{\plaindom{Conf}}
-\newcommand{\Instrs}{\plaindom{Instructions}}
-\newcommand{\nats}{\mathbb{N}}
-\newcommand{\ints}{\mathbb{Z}}
-\newcommand{\Perms}{\plaindom{Perm}}
-\newcommand{\Globals}{\plaindom{Global}}
-
-\newcommand{\Rel}{\plaindom{Rel}}
-\newcommand{\Rels}{\plaindom{Rels}}
-\newcommand{\States}{\plaindom{RState}}
-\newcommand{\RegionNames}{\plaindom{RegionName}}
-\newcommand{\Regions}{\plaindom{Region}}
-\newcommand{\Reg}{\plaindom{Reg}}
-\newcommand{\Worlds}{\plaindom{World}}
-\newcommand{\Wor}{\plaindom{Wor}}
-\newcommand{\Worwk}{\Wor_{\futurewk}}
-\newcommand{\Worstr}{\Wor_{\futurestr}}
-\newcommand{\xiwk}{\xi_{\var{wk}}}
-\newcommand{\xistr}{\xi_{\var{str}}}
-\newcommand{\StorePred}{\plaindom{MemSegPred}}
-\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}
-\newcommand{\DCPred}[1]{\plaindom{P}^\downarrow(#1)}
-
-\newcommand{\Views}{\plaindom{View}}
-
-% LR
-\newcommand{\intr}[2]{\mathcal{#1}}
-\newcommand{\valueintr}[1]{\intr{V}{#1}}
-\newcommand{\exprintr}[1]{\intr{E}{#1}}
-\newcommand{\contintr}[1]{\intr{K}{#1}}
-\newcommand{\regintr}[1]{\intr{R}{#1}}
-\newcommand{\stdvr}{\valueintr{\asmType}}
-\newcommand{\stder}{\exprintr{\asmType}}
-\newcommand{\stdrr}{\regintr{\asmType}}
-\newcommand{\stdkr}{\contintr{\asmType}}
-\newcommand{\observations}{\mathcal{O}}
-\newcommand{\npair}[2][n]{\left(#1,#2 \right)}
-\newcommand{\npairP}[2][n]{(#1,#2)}
-
-% Reference register/memory
-\newcommand{\refreg}[1]{#1}
-\newcommand{\refheap}[1]{#1}
-
-% Instructions
-% No arguments
-\newcommand{\zinstr}[1]{\mathtt{#1}}
-\newcommand{\fail}{\zinstr{fail}}
-\newcommand{\halt}{\zinstr{halt}}
-% One argument
-\newcommand{\oneinstr}[2]{
-  \ifthenelse{\equal{#2}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2}
-}
-\newcommand{\jmp}[1]{\oneinstr{jmp}{#1}}
-% Two arguments
-\newcommand{\twoinstr}[3]{
-  \ifthenelse{\equal{#2#3}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3}
-}
-\newcommand{\restricttwo}[2]{\twoinstr{restrict}{#1}{#2}}
-\newcommand{\jnz}[2]{\twoinstr{jnz}{#1}{#2}}
-\newcommand{\isptr}[2]{\twoinstr{isptr}{#1}{#2}}
-\newcommand{\geta}[2]{\twoinstr{geta}{#1}{#2}}
-\newcommand{\getb}[2]{\twoinstr{getb}{#1}{#2}}
-\newcommand{\gete}[2]{\twoinstr{gete}{#1}{#2}}
-\newcommand{\getp}[2]{\twoinstr{getp}{#1}{#2}}
-\newcommand{\getl}[2]{\twoinstr{getl}{#1}{#2}}
-\newcommand{\move}[2]{\twoinstr{move}{#1}{#2}}
-\newcommand{\store}[2]{\twoinstr{store}{#1}{#2}}
-\newcommand{\load}[2]{\twoinstr{load}{#1}{#2}}
-\newcommand{\lea}[2]{\twoinstr{lea}{#1}{#2}}
-% Three arguments
-\newcommand{\threeinstr}[4]{
-  \ifthenelse{\equal{#2#3#4}{}}
-  {\zinstr{#1}}
-  {\zinstr{#1} \; #2 \; #3 \; #4}
-}
-\newcommand{\restrict}[3]{\threeinstr{restrict}{#1}{#2}{#3}}
-\newcommand{\subseg}[3]{\threeinstr{subseg}{#1}{#2}{#3}}
-\newcommand{\plus}[3]{\threeinstr{plus}{#1}{#2}{#3}}
-\newcommand{\minus}[3]{\threeinstr{minus}{#1}{#2}{#3}}
-\newcommand{\lt}[3]{\threeinstr{lt}{#1}{#2}{#3}}
-
-% Permissions
-\newcommand{\plainperm}[1]{\textsc{#1}}
-\newcommand{\noperm}{\plainperm{o}}
-\newcommand{\readonly}{\plainperm{ro}}
-\newcommand{\readwrite}{\plainperm{rw}}
-\newcommand{\exec}{\plainperm{rx}}
-\newcommand{\entry}{\plainperm{e}}
-\newcommand{\rwx}{\plainperm{rwx}}
-% PWL permissions
-\newcommand{\readwritel}{\plainperm{rwl}}
-\newcommand{\rwl}{\readwritel}
-\newcommand{\rwlx}{\plainperm{rwlx}}
-
-% Global/local
-\newcommand{\plainlocality}[1]{\mathrm{#1}}
-\newcommand{\local}{\plainlocality{local}}
-\newcommand{\glob}{\plainlocality{global}}
-
-\newcommand{\localityReg}{\var{localityReg}}
-\newcommand{\localReg}{\var{localReg}}
-\newcommand{\globalReg}{\var{globalReg}}
-
-% Views
-\newcommand{\plainview}[1]{\mathrm{#1}}
-\newcommand{\perma}{\plainview{perm}}
-\newcommand{\temp}{\plainview{temp}}
-\newcommand{\revoked}{\plainview{revoked}}
-
-% OP sem
-\newcommand{\diverge}[1][n]{\not\Downarrow_{#1}}
-\newcommand{\step}[1][]{\rightarrow_{#1}}
-
-% Conv defs
-\newcommand{\linksto}[4]{\ensuremath{#1} \text{ links } \ensuremath{#2} \text{ as } \ensuremath{#3} \text{ to } \ensuremath{#4}}
-\newcommand{\lookingat}[3]{\ensuremath{#1} \text{ is looking at } \ensuremath{#2} \text{ followed by } \ensuremath{#3}}
-\newcommand{\pointstostack}[3]{\ensuremath{#1} \text{ points to stack with } \ensuremath{#2} \text{ used and } \ensuremath{#3} \text{ unused}}
-\newcommand{\nonlocal}[1]{\ensuremath{#1} \text{ is non-local}}
-
-% Macros
-\newcommand{\scall}[3]{\mathtt{scall} \; #1([#2],[#3])}
-
-% Text
-\newcommand{\tpwl}{permit-write-local standard region}
-\newcommand{\tnwl}{no-write-local standard region}
-
-\newcommand{\isdef}{\mathrel{\overset{\makebox[0pt]{\mbox{\normalfont\tiny\sffamily def}}}{=}}}
-\newcommand\bnfdef{\mathrel{::=}}
-
-\definecolor{OliveGreen}{RGB}{60,128,49}
-\definecolor{Blue}{RGB}{92,26,143}
-\newenvironment{toplas}%{}{}
-    {\color{OliveGreen}}{}
-\newenvironment{toplassug}
-    {\begin{comment}}{\end{comment}}
-
-\newcommand{\itoplas}[1]%{#1}
-    {{\color{OliveGreen} #1}}
-\newcommand{\itoplassug}[1]{}
-%    {{\color{Blue} #1}}
-
-\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
-% domi: IEEEtran.cls sets font size to 10bp, apparently.
-% not sure what the difference is with 10pt.
-\DeclareMathSizes{10bp}{8}{7}{6}
-
-\begin{document}
-
-%% Title information
-%\title{Reasoning About a Capability Machine with Local Capabilities}         %% [Short Title] is optional;
-\title{Reasoning About a Machine with Local Capabilities}        %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-%\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-%\subtitle{Provably Safe Stack and Return Pointer Management (without OS Support)}                     %% \subtitle is optional
-\subtitle{Provably Safe Stack and Return Pointer Management}                     %% \subtitle is optional
-%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-\bibliographystyle{ACM-Reference-Format}
-\citestyle{acmauthoryear}
-
-% Authors and affiliation
-%% Author with single affiliation.
-\author{Lau~Skorstengaard}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-  \department{Department of Computer Science}              %% \department is recommended
-  \institution{Aarhus University}            %% \institution is required
-  \streetaddress{\AA{}bogade 34}
-  \city{Aarhus}
-  \postcode{8210}
-  \country{Denmark}
-}
-\email{lau@cs.au.dk}          %% \email is recommended
-
-\author{Dominique~Devriese}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-  \department{iMinds-DistriNet}              %% \department is recommended
-  \institution{KU Leuven}            %% \institution is required
-  \streetaddress{Celestijnenlaan 200 A}
-  \city{Leuven}
-  \postcode{3001}
-  \country{Belgium}
-}
-\email{dominique.devriese@cs.kuleuven.be}          %% \email is recommended
-
-\author{Lars~Birkedal}
-%\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-%  \position{Position1}
-  \department{Department of Computer Science}              %% \department is recommended
-  \institution{Aarhus University}            %% \institution is required
-  \streetaddress{\AA{}bogade 34}
-  \city{Aarhus}
-  \postcode{8210}
-  \country{Denmark}
-}
-\email{birkedal@cs.au.dk}          %% \email is recommended
-% \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
-% \institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
-% \and imec-DistriNet, KU Leuven \\ \email{dominique.devriese@cs.kuleuven.be}}
-
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986.10002990</concept_id>
-<concept_desc>Security and privacy~Logic and verification</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10011007.10010940.10010992.10010998.10010999</concept_id>
-<concept_desc>Software and its engineering~Software verification</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-\ccsdesc[500]{Security and privacy~Logic and verification}
-\ccsdesc[500]{Software and its engineering~Software verification}
-
-\begin{abstract}
-  Capability machines provide security guarantees at machine level which makes them an interesting target for secure compilation schemes that provably enforce properties such as control-flow correctness and encapsulation of local state.
-  We provide a formalization of a representative capability machine with local capabilities and study a novel calling convention.
-  We provide a logical relation that semantically captures the guarantees provided by the hardware (a form of capability safety) and use it to prove control-flow correctness and encapsulation of local state.
-  The logical relation is not specific to our calling convention and can be used to reason about arbitrary programs.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Compromising software security is often based on attacks that break programming language properties relied upon by software authors such as control-flow correctness, local-state encapsulation, etc.
-Commodity processors offer little support for defending against such attacks: they offer security primitives with only coarse-grained memory protection and limited compartmentalization scalability.
-As a result, defenses against attacks on control-flow correctness and local-state encapsulation are either limited to mitigation of only certain common forms of attacks (leading to an attack-defense arms race \citep{Szekeres_sok_2013}) and/or rely on techniques like machine code rewriting~\citep{wahbe_efficient_1993,abadi_control-flow_2005}, machine code verification~\citep{morrisett_system_1999}, virtual machines with a native stack~\citep{lindholm_java_2014} or randomization~\citep{forrest_building_1997}.
-The latter techniques essentially emulate protection techniques on existing hardware at the cost of performance, system complexity, and/or security.
-
-\emph{Capability machines} are a type of processors that remediate these limitations with a better security model at the hardware level.
-They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that have recently received renewed interest.
-in particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
-Capability machines tag every word (in the register file and in memory) to enforce a strict separation between numbers and capabilities (a kind of pointers that carry authority).
-Memory capabilities carry the authority to read and/or write to a range of memory locations.
-There is also a form of \emph{object capabilities}, which represent the authority to invoke a piece of code without exposing the code's encapsulated private state (e.g., the M-Machine's enter capabilities (described in Section~\ref{sec:capab-mach-with}) or CHERI's sealed code/data pairs).
-
-Unlike commodity processors, capability machines lend themselves well to
-enforcing local-state encapsulation. Potentially, they will enable compilation
-schemes that enforce this property in an efficient but also 100\% watertight way
-(ideally evidenced by a mathematical proof, guaranteeing that we do not end up
-in a new attack-defense arms race). However, a lot needs to happen before we get
-there. For example, it is far from trivial to devise a compilation scheme
-adapted to the details of a specific source language's notion of encapsulation
-(e.g., private member variables in OO languages often behave quite differently
-than private state in ML-like languages). And even if such a scheme were
-defined, a formal proof depends on a formalization of the encapsulation provided
-by the capability machine at hand.
-
-A similar problem is the enforcement of control-flow correctness on capability machines.
-An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}: the standard contiguous C stack is split into a central, trusted stack and disjoint, private, per-compartment stacks.
-The trusted stack is managed by trusted call and return instructions.
-To prevent illegal use of stack references, the approach relies on \emph{local capabilities}, a type of capabilities offered by CHERI to \emph{temporarily} relinquish authority, namely for the duration of a function invocation whereafter the capability can be revoked.
-However, details are scarce (how does it work precisely?
-what features are supported?)
-and a lot remains to be investigated (e.g., combining disjoint stacks with cross-domain function pointers seems like it will scale poorly to large numbers of components).
-Finally, there is no argument that the approach is watertight and it is not even clear what security property is targeted exactly.
-
-\dominique{I think we should include a comparison with the conference paper, and explain what the delta is for this journal version.}
-In this paper, we make two main contributions: (1) an alternative calling
-convention that uses local capabilities to enforce stack frame encapsulation and
-well-bracketed control flow, and (2) perhaps more importantly, we adapt and
-apply the well-studied techniques of step-indexed Kripke logical relations for
-reasoning about code on a representative capability machine with local
-capabilities in general and correctness and security of the calling convention
-in particular. More specifically, we make the following contributions:
-\begin{itemize}
-\item We formalize a simple but representative capability machine featuring local
-  capabilities and its operational semantics
-  (\sectionname~\ref{sec:capab-mach-with}).
-\item We define a novel calling convention enforcing control-flow correctness and
-  encapsulation of stack frames (\sectionname~\ref{sec:stack-and-return-pointer}). It
-  relies solely on local capabilities and does not require OS support (like a
-  trusted stack or call/return instructions).
-  %It also does not use separate per-component stacks but a single, fairly standard, contiguous stack.
-  %domi: don't mention this here. I think some CSF reviewers misunderstood that we were claiming this as a strength of the approach
-  It supports higher-order cross-component calls (e.g., cross-component function
-  pointers) and can be efficient assuming only one additional piece of processor
-  support (w.r.t. CHERI): an efficient instruction for clearing a range of memory.
-\item We present a novel step-indexed Kripke logical relation for reasoning
-  about programs on the capability machine. It is an untyped logical relation,
-  inspired by previous work on object capabilities \citep{Devriese:2016ObjCap}.
-  We prove an analogue of the standard fundamental theorem of logical relations
-  --- to the best of our knowledge, our theorem is % by far
-  the most general and powerful formulation of the formal guarantees offered by
-  a capability machine (a form of capability
-  safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}), including the specific
-  guarantees offered for local capabilities. It is very general and not tied to
-  our calling convention or a specific way of using the system's capabilities.
-  We are the first to apply these techniques for reasoning about capability
-  machines and we believe they will prove useful for many other purposes than
-  our calling convention.
-\item We introduce two novel technical ideas in the unary, step-indexed Kripke
-  logical relation used to formulate the above theorem: the use of a
-  \emph{single} orthogonal closure (rather than the earlier used biorthogonal
-  closure) and a variant of \citet{Dreyer:jfp12}'s public and private future
-  worlds~\citep{Dreyer:jfp12} to express the special nature of local
-  capabilities. The logical relation and the fundamental theorem expressing
-  capability safety are presented in \sectionname~\ref{sec:logical-relation}.
-\item We demonstrate our results by applying them to challenging examples,
-  specifically constructed to demonstrate local-state encapsulation and
-  control-flow correctness guarantees in the presence of cross-component
-  function pointers (\sectionname~\ref{sec:examples}). The examples demonstrate both
-  the power of our formulation of capability safety and our calling convention.
-\end{itemize}
-%\lau{I have left out the discussion/future work, related work and conclussion sections from the outline given with the contribution.
-Do you feel like there should be a dedicated outline instead and if not should we mention the last sections (I don't feel it contributes anything to mention them).}
-\begin{toplas}
-  This paper is an extension of the published conference paper~\citet{Skorstengaard:esop18}.
-  We have made improvements to readability and completeness throughout the paper.
-  For instance, we have added an introduction to \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
-  We highlight the following changes:
-  \begin{itemize}
-  \item We have added proof sketches for the \nameref{thm:ftlr} and for the correctness lemma for the awkward example
-    (Lemma~\ref{lem:correctness-g1}).
-  \item We have added figures that illustrate central parts of the calling convention.
-  \item A section about malloc has been added.
-    Specifically, we provide the specification for the malloc used in the examples of the paper.
-  \item A section about macro instructions with descriptions of all the macros and the implementation of \texttt{scall} has been added.
-  \item We have added a section on reasoning about programs that run on a capability machine.
-    This section explains how one reasons about common scenarios that arise in programs on a capability machine, and, in particular, how the logical relation is used.
-    It also introduces a number of lemmas that prove recurring bits once and for all.
-  \item We have expanded on the explanation of the awkward example. 
-  \item Details previously found only in the technical appendix~\citep{technical_appendix} have been moved to the paper.
-\end{itemize}
-\end{toplas}
-We have written a technical appendix~\citep{technical_appendix} which contains
-additional details and proofs left out from this paper.
-
-% % Possible high-level intros:
-% % 0) Arms race compilers
-% %
-% % 1) Security guarantees vs obstacles / arms race
-% \begin{comment}
-%   Security on processors is today subject to an arms race between
-%   processor manufacturers and hackers. On the hacker side, a constant
-%   effort is put into exploiting current processor designs and
-%   circumventing the security measures set in place to prevent known
-%   exploits. On the processor manufacturer side, new security measures
-%   are put in place to prevent new exploits. The arms race would come
-%   to a grinding halt if the security meassures provided more than
-%   obstacles and actually provided low-level security guarantees. One
-%   kind of low-level machines that provide low-level security is a
-%   capability machine.
-% \end{comment}
-% % 2) Fully abstract compilation
-% High-level languages provide guarantees such as encapsulation of local
-% state and well-bracketedness of function calls. These guarantees are
-% often relied upon when reasoning about the correctness and security of
-% a program. High-level programs are compiled to machine code, so in
-% order to rely on the correctness and security guarantees provided by
-% the high-level language, it needs to be proven that the translation
-% preserves these properties. Translated programs often interact with
-% machine code that is not translated from the same high-level language,
-% so in order to preserve the guarantees from the high-level languages,
-% the low-level machine needs to provide some kind of security
-% guarantees. In other words, a target of secure compilation needs to
-% provide some security guarantees. Modern processors provide many
-% security meassures, but very little in terms of guarantees. A machine
-% that provides some means of security guarantees is a capability machine.
-
-% % 3) Memory access-control
-% % I don't like this angle and can't find a good way to sell it.
-
-% % 4) History
-% % Proposals throughout history.
-% % Lack of formal account of properties-guarantees.
-
-% %%% a)
-% % Current low-level protection coarse grained memory protection or
-% % properties of high-level languages are not really enforced?
-
-% % Capability machines offer fine-grained memory protection
-% % Too technical
-% A capability machine is a low-level machine that offers fine-grained
-% memory protection.
-% % Capability machines and unforgeable tokens for memory access. 
-% This is done by introducing unforgeable tokens to
-% the machine. These tokens are called capabillities and grant a kind of
-% authority and a range of authority. 
-% % Dynamic checks encapsulation, formal model
-
-%  We will refer to the capabilities
-% that grant a combination of read, write, or execute permission as memory
-% capabilities. 
-%  Capabilities with the necessary permissions must be
-% presented every time an instruction that manipulates the memory is
-% executed.
-
-% Memory capabilities provide memory protection, but they do not provide
-% a way to set up security domains. When you jump to an untrusted piece
-% of code that is supposed to jump back to you, then you have to either
-% revoke all your capabilities or pass to the piece of code you do not
-% trust. This short coming can be overcome by adding something like the
-% enter capability from the M-Machine\todo{add reference}. The enter
-% capability is an opaque capability which can only be used for a
-% jump. When jumped to it grants read and execute permission which
-% allows the code that is now executing to read capabilities stored in
-% memory. 
-% The CHERI processor's ccall achieves a similar property\todo{Add reference}.
-
-% Capabilities are irrevocable, so when we pass a capability has to
-% an untrusted program, then we have to assume that the program keeps it
-% around indefenitely. This means that we cannot reuse the piece of
-% memory that the capability governs essentially creating a memory
-% leak. The CHERI processor has a special kind of capabilities called
-% \emph{local capabilities}\todo{reference?}\ that introduces a simple
-% kind of temporal information control which allows for a simple kind of
-% capability revocation. This is achieved by adding a tag to every
-% capability that marks it as either local or global. Local capabilities
-% can only be written through a capability with a new permission called
-% \emph{permit write local}.
-% % More about local capabilities.
-
-% With just memory capabilities, enter capabailities, and local
-% capabilities, the capability machine is very powerful and can
-% enforce properties that high-level languages promise. In most
-% high-level languages, we expect local state to be private and
-% encapsulated. The memory capabilities make sure that local state
-% cannot be accessed without a capability to do so. With the enter
-% capabaility, we can regain control of local state after passing
-% control to an untrusted piece of code which gives us encapsulation
-% when control is passed to an untrusted program.
-
-% Many high-level languages guarantee that function function calls are
-% well-bracketed. This can easily be achieved by having a trusted call
-% stack. On a capability machine, it is possible to enforce well
-% barcketedness without a trusted stack. When ensuring well-bracketed
-% the two main challenges calls are 1) preventing storage of return
-% pointers and 2) encapsulation of the local stack frame. If we don't
-% have the first property, then an adversary can store the return
-% capability in a call and jump to it in a later call. If the local
-% stack frame is not properly encapsulated, then an adversary can break
-% the well-bracketedness by jumping to a different calls return
-% capability.
-
-% In order to deal with the two challenges, we make sure that all return
-% capabilities are local and that there are no global capabilities with
-% permit write local permission. This makes sure that there is no way to
-% save a return capability in persistent local state.\lau{this does not really make
-% sense as the local stack frame can be seen as local state} It is,
-% however, to limiting to not provide any means to store the return
-% capabilities as a piece of code may need to store a return capability
-% before jumping to an adversary. It is therefore needed to provide a
-% local capability with permit write local permission which can be used
-% as a stack. The stack should now be the only place one can store the
-% return capability, but it should also be the only place one can store
-% the stack capability.
-
-% Using local return capabilities and putting a stack abstraction on a
-% local capability permit write local capability is a good start, but as
-% it turns out, it is not wnough. An adversary could do the following 
-% \begin{enumerate}
-% \item In a first invocation, fill the stack frame with copies of the
-%   current stack capability.
-% \item In a later call, the adversary may be able to load the stack
-%   capability from the previous call.
-% \item If we have used a larger part of the stack than in the first
-%   call, then the adversary has access to part of our local stack frame.
-% \end{enumerate}
-
-% The above attack shows that it is tricky to ensure
-% well-bracketedness. It raises the question whether it is possible to
-% do so it is completely watertight. The attack utilized the stack, so
-% it is necessary to have a trusted stack or can we do without? In
-% general it raises the question, how do we reason about capability
-% machines and local capabilities, so we are sure that what ever calling
-% scheme we come up with actually provably ensures well-bracketed calls.
-
-% In the following paper, we present the following contributions
-% \begin{enumerate}
-% \item 
-% \end{enumerate}
-
-%%% b)
-
-% Examples of capability machines?
-
-% Enter capabilities
-
-% Local capabilities
-
-% High-level languages often provide guarantees such as encapsulation of private state and well-bracketedness of calls - often not clear how well this is enforced.
-Just enforced when interacting with other programs written in the same high-level language or is it also guaranteed when interacting with assembly programs.
-
-% High-level programs compiled to assembly can ensure these properties.
-
-%
-
-% \subsection{Contributions}
-
-% \begin{itemize}
-% \item Formal model of a simple but representative capability machine with local capabilities
-% \item Detailed study of how to do safe stack and return pointer management in this setting, taking into account:
-% \begin{itemize}
-% \item untrusted adversary
-% \item higher-order code: callbacks passed to and received from the adversary
-% \item efficient stack management
-% \item no OS support
-% \end{itemize}
-% \item Logical relation for reasoning about code in this capability machine
-% \item Fundamental theorem that expresses the guarantees offered by a capability machine for untrusted code
-% \item Technical contributions: reuse existing ideas in a new way:
-% \begin{itemize}
-% \item replace biorthogonal closure by a single orthogonal closure
-% because assembly languages remove the distinction between the
-% continuation and the arguments
-% \end{itemize}
-% \end{itemize}
-% LB: I think we should have some discussion of the flexibility /
-%   strength of our LR.  The LR defines which computations we think of
-%   as well-behaved.   
-%  We should give simple / trivial examples of
-%   code not in the LR. But we may also want to emphasize that the LR is 
-%   not too restricting, e.g., it does not enforce a certain calling
-%   scheme (examplied by Example \verb!f1! which
-%   does not use the stack and the following examples which do use the
-%   stack). This ties in to the discussion of biorthogonality, see the
-%   following paragraphs.
-
-% LB: we need to relate this carefully to Hur-Dreyer, who used
-%       biorthogonality even though they also worked with an assembly
-%       language. (If I recall correctly, they assumed some properties
-%       of the low level language, which is why the biortho was the
-%       right ting, but we need to check.)
-
-% LS: It is my impression that they were interested in the relation
-% between high-level and low-level programs. They are therefore in
-% particular interested in low-level programs compiled from high-level
-% programs which means that the continuation is always invoked in the 
-% same way. Our realization was that anything we pass as an argument 
-% in a "call" can be used as a continuation, so the continuation 
-% relation was redundant. Even though we do try to give our programs
-% some structure using the macros, our logical relation is strong enough
-% to handle unstructured programs as well.
-
-% DD: I think the above comment by LS is correct: the reason why the biortho is
-% the right thing for Hur-Dreyer is that their LR only accepts well-behaved
-% programs which treat the return pointer as a continuation (i.e. don't store
-% it), while ours is more general.
-
-% \begin{itemize}
-% \item STSs with public/private transitions for dealing with local capabilities: play the same role as before, but in different places of the LR.
-% \end{itemize}
-% \begin{itemize}
-% \item Demonstrate all of this on several examples:
-% \begin{itemize}
-% \item security examples
-% \item the most challenging examples from existing literature on reasoning about well-bracketed control flow in lambda calculi
-% \item some compartmentalization result?
-% \item whatever else we do..
-% \end{itemize}
-% \end{itemize}
-
-\section{A Capability Machine with Local Capabilities}
-\label{sec:capab-mach-with}
-In this paper, we work with a formalization of a capability machine with all the characteristics of real capability machines as well as local capabilities much like CHERI's.
-Otherwise, it is kept as simple as possible.
-It is inspired by both the M-Machine~\citep{Carter:1994:HSF:195473.195579} and CHERI~\citep{Watson2015Cheri}.
-For simplicity, we assume an
-infinite address space and unbounded integers (see \sectionname~\ref{sec:discussion} for a discussion of these assumptions).
-
-We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
-We assume an infinite set of addresses $\Addrs$ and define machine words as either integers or capabilities of the form $((\perm,\gl),\var{base},\var{end},\addr)$.
-Such a capability represents the authority to execute permissions $\perm$ on the memory range $[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag $\gl$ indicating whether the capability is global or local.
-On a capability machine, there is no notion of pointers other than capabilities, so we will use the terms interchangeably.
-The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$), and read/write/execute ($\rwx$) permissions.
-Additionally, there are three special permissions: read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter ($\entry$), which we will explain below.
-\itoplas{The orderings of permissions and locality are displayed in \figurename~\ref{fig:perm-hier}.
-  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory which is why write-local is above write in the permission hierarchy.
-  We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
-
-\begin{wrapfigure}{R}{.4\textwidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (7) {$\rwlx$};
-    \node[main node] (8) [below left of=7] {$\readwritel$};
-    \node[main node] (1) [below right of=7] {$\rwx$};
-    \node[main node] (2) [below right of=1] {$\exec$};
-    \node[main node] (3) [below right of=2] {$\entry$};
-    \node[main node] (4) [below left of=1] {$\readwrite$};
-    \node[main node] (5) [below right of=4] {$\readonly$};
-    \node[main node] (6) [below right of=5] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (7) edge (8)
-    (7) edge (1)
-    (8) edge (4)
-    (1) edge (2)
-    (2) edge (3)
-    (2) edge (5)
-    (3) edge (6)
-    (1) edge (4)
-    (4) edge (5)
-    (5) edge (6);
-
-    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
-    \node[main node] (2l) [below of=1l] {$\local$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (1l) edge (2l);
-  \end{tikzpicture}
-  \caption{Permission and locality hierarchy.}
-  \label{fig:perm-hier}
-\end{wrapfigure}
-
-We assume a finite set of register names $\RegName$.
-We define register files $\reg$ and memories $\ms$ as functions mapping register names resp.\ addresses to words.
-The state of the entire machine is represented as a configuration that is either a running state $\Phi \in \ExecConfs$ containing a memory and a register file or a failed or halted state where the latter is paired with the final state of memory.
-
-% We also define memory segments $\ms$:
-% parts of memory represented as partial functions mapping addresses to words.
-
-%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
-
-The machine's instruction set is rather basic.
-Instructions $i$ include relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$), and move ($\move{}{}$, copies words between registers) instructions.
-Also familiar are load and store instructions for reading from and writing to memory ($\load{}{}$ and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and $\minus{}{}{}$, operating only on numbers).
-There are three instructions for modifying capabilities: $\lea{}{}$ (modifies the current address), $\restricttwo{}{}$ (modifies the permission and local/global tag), and $\subseg{}{}{}$ (modifies the range of a capability).
-Importantly, these instructions take care that the resulting capability always carries less authority than the original (e.g.\ $\restricttwo{}{}$ will only weaken a permission according to the hierarchy in \figurename~\ref{fig:perm-hier}).
-Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and $\geta{}{}$ provide access to a capability's permissions, local/global tag, base, end and current address, respectively.
-
-\begin{figure}[b]
-\[\arraycolsep=1.4pt
-  \begin{array}{rlrcl}
-    \addr   &\in& \Addrs &\isdef& \nats \\
-    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid\exec\mid \entry\mid \rwx\mid \rwlx  \\
-    \gl&\in&\Globals & ::= &\glob \mid \local \\
-    c &\in& \Caps  &\isdef & \left\{
-                         ((\perm,\gl),\start,\addrend,\addr)\mid
-                         \start,\addr\in\Addrs,\addrend \in
-                         \Addrs\cup \{\infty\}\right\}\\
-    w &\in&\Words &\isdef& \ints + \Caps \\
-    r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
-    \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
-    m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
-    \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
-    \ms  &\in & \MemSegments &\isdef& \Addrs \parfun \Words \\
-    &&\Confs &\isdef& \ExecConfs + \{\failed \} + \{\halted\} \times \Heaps
-  \end{array}
-\]
-  \begin{equation*}
-  \begin{array}{rcl}
-    \rv    &\in& \ints + \RegName \\
-    i      &::=& 
-                 \jmp{\lv} \mid 
-                 \jnz{\lv}{\rv} \mid
-                 \move{\lv}{\rv} \mid 
-                 \load{\lv}{\hv} \mid 
-                 \store{\hv}{\rv} \mid
-                 \lt{\lv}{\rv}{\rv} \mid
-                 \plus{\lv}{\rv}{\rv} \mid 
-                 \minus{\lv}{\rv}{\rv} \mid \\
-          &   &  \lea{\lv}{\rv} \mid 
-                 \restricttwo{\lv}{\rv} \mid
-                 \subseg{\lv}{\rv}{\rv} \mid  
-                 \isptr{\lv}{\rv} \mid 
-                 \getl{\lv}{\lv} \mid
-                 \getp{\lv}{\lv} \mid
-                 \getb{\lv}{\lv} \mid\\
-          &   &  \gete{\lv}{\lv} \mid 
-                 \geta{\lv}{\lv} \mid 
-                 \fail \mid
-                 \halt 
-  \end{array}
-\end{equation*}
-  \caption{The syntax of our capability machine assembly language.}
-  \label{fig:syntax}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \begin{equation*}
-    \Phi  \rightarrow
-    \begin{cases}
-      \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
-      \begin{array}[t]{l}
-        \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$}\\ 
-        \text{\;\;and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }
-      \end{array}
-\\
-      \failed                                 & \text{otherwise}
-    \end{cases}
-  \end{equation*}
-  \begin{align*}
-    \stdUpdatePc{\Phi} &=
-                         \begin{cases}
-                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
-                           \begin{array}[t]{l}
-                             \text{if $\memreg(\pcreg) = \stdcap$}\\
-                             \text{\;\;and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
-                           \end{array}
-\\
-                             \failed & \text{otherwise}
-                         \end{cases}
-  \end{align*}
-  \begin{tabular}{|c|p{3.4cm}|>{\raggedright\arraybackslash}p{7.3cm}|}
-    \hline
-    $i$&$\sem{i}(\Phi)$&Conditions\\
-    \hline 
-    $\fail$&$\failed$&\\
-    \hline
-    $\halt$&$(\halted,\memheap)$&\\
-    \hline
-    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{w})$&$r_2 \in \Regs \Rightarrow w = \memreg(r_2)$ and $r_2 \in \ints \Rightarrow w = r_2$\\
-    \hline
-    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and  $w = \memheap(\addr)$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
-    \hline
-    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$ and $(\perm',g') = \decodePermPair{\memreg(r_2)}$
-                                                                              and $(\perm',g') \sqsubseteq (\perm,g)$ and  $w =((\perm',g'),\start,\addrend,\addr)$\\
-    \hline
-    \itoplas{$\subseg{\refreg{r_1}}{r_2}{r_3}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{w}}}$} &  \itoplas{$\memreg(r_1) = \stdcap$ and for $i \in \{2,3\}$ $n_i = \memreg(\rv_i)$ and $n_2 \in \nats$ and $\start \leq n_2$ and $n_3 \leq \addrend$ where either $n_3 \in \nats$ or ($n_3=\inftyend$ and $\addrend = \infty$) and $\perm \neq \entry$ and $w = ((\perm,\gl),n_2,n_3,\addr)$}\\
-    \hline
-    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
-    \hline
-    \itoplas{$\lea{\refreg{r_1}}{r_2}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{c}}}$} &\itoplas{$\memreg(r_1) = \stdcap$ and $n = \memreg(r_2)$ and $n \in \ints $ and $\perm \neq \entry$ and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$} \\
-    \hline
-    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
-                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
-    \hline
-    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
-                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
-    \hline
-    \multicolumn{3}{|c|}{$\cdots$}\\
-    \hline
-    \_&$\failed$&otherwise\\
-    \hline
-  \end{tabular}
-  \caption{An excerpt from the operational semantics.}
-  \label{fig:op-sem}
-\end{figure}
-
-\begin{comment}
-  \begin{figure}
-    \centering
-    \begin{mathpar}
-      \Phi \rightarrow
-      \begin{cases}
-        \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
-        \begin{array}[t]{l}
-          \text{if $\memreg(\pcreg) = \stdcap$}\\
-          \;\;\text{and $\start \leq \addr \leq \addrend$ and $\memheap(\addr) = n$}\\
-          \;\;\text{and $\perm \in \{ \exec,\rwx, \rwlx \}$} 
-        \end{array}\\
-        \failed                & \text{otherwise}
-      \end{cases}
-      \and \updatePcPerm{w} =
-      \begin{cases}
-        ((\exec,\gl),\start,\addrend,\addr) & \text{\footnotesize{if $w = ((\entry,\gl),\start,\addrend,\addr)$}}\\
-        w & \text{otherwise}
-      \end{cases}\\
-      \and \stdUpdatePc{\Phi} =
-      \begin{cases}
-        \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
-        \begin{array}[t]{l}
-          \text{if $\memreg(\pcreg) = \stdcap$}\\ 
-          \;\;\text{and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
-        \end{array}\\
-        \failed & \text{otherwise}
-      \end{cases}
-      \and \inferrule{ } { \sem{\fail}(\Phi) = \failed } \and
-      \inferrule{ } { \sem{\halt}(\Phi) = (\halted,\memheap) } \and
-      \inferrule{ } { \sem{\move{\refreg{r_1}}{r_2}}(\Phi) =
-        \stdUpdatePc{}(\updateReg{r_1}{\memreg(r_2)}) } \and
-      \inferrule{ \memreg(r_2) = \stdcap \\
-        \perm \in \{ \rwx, \rwlx, \exec, \readwrite,
-        \readwritel, \readonly \}\\
-        \start \leq \addr \leq \addrend } {
-        \sem{\load{\refreg{r_1}}{\refheap{r_2}}}(\Phi) =
-        \updateReg{r_1}{\memheap(\addr)}} \and
-      \inferrule{ \memreg(r_2) = \stdcap[\permp]\\
-        \permp' = \decodePermPair{\memreg(r_2)}\\
-        \permp'\sqsubseteq \permp } {
-        \sem{\restricttwo{\refreg{r_1}}{r_2}}(\Phi) = \\
-        \stdUpdatePc{}(\updateReg{r_1}{(\permp',\start,\addrend,\addr)})
-      } \and \inferrule{ \memreg(r_2) = ((\_,\_),\_,\_,\addr) } {
-        \sem{\geta{\refreg{r_1}}{\refreg{r_2}}}(\Phi) =
-        \stdUpdatePc{\updateReg{r_1}{\addr}}} \and \inferrule{
-        \var{newPc} = \updatePcPerm{\memreg(\lv)} } {
-        \sem{\jmp{\lv}}(\Phi) = \updateReg{\pcreg}{\var{newPc}} } \and
-      \inferrule{ \memreg(r_1) = \stdcap\\
-        \start \leq \addr \leq \addrend\\
-        \perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}\\
-        \var{w} = \memreg(r_2) \\
-        \var{w} = ((\_,\local),\_,\_,\_)\Rightarrow\perm \in
-        \{\rwlx,\readwritel \} } {
-        \sem{\store{\refheap{r_1}}{\refreg{r_2}}}(\Phi) =
-        \stdUpdatePc{\updateHeap{\addr}{\var{w}}} } \and
-    \end{mathpar}
-    \caption{An excerpt from the operational semantics.}
-    \label{fig:op-sem}
-  \end{figure}
-\end{comment}
-\figurename~\ref{fig:op-sem}
-%\itoplassug{Suggestion: extend this figure with more (all?) of the operational semantics.
-In general, should we add all the missing definitions and make this document closer to being self-contained.}
-shows the operational semantics for a few instructions representative of the machine.
-Essentially, a configuration $\Phi$ either decodes and executes the instruction pointed to by $\memreg(\pcreg)$ if it is an executable capability with its address in the valid range; otherwise it fails.
-The table in the figure shows for instructions $i$ the result of executing them in configuration $\Phi$.
-The instructions $\fail$ and $\halt$ obviously fail and halt respectively.
-$\move{}{}$ simply modifies the register file as requested and updates the $\pcreg$ to the next instruction using the meta-function $\stdUpdatePc{}$.
-
-\dominique{Follow order of figure in discussion?}
-The $\load{}{}$ instruction loads the contents of the requested memory location
-into a register, but only if the capability has appropriate authority (i.e.\
-read permission and an appropriate range). The $\restricttwo{}{}$ instruction updates a
-capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original according to the permission hierarchy in \figurename~\ref{fig:perm-hier}.
-%\itoplassug{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
-%\itoplassug{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
-% Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
-The $\subseg{}{}{}$ instruction reduces the range of authority of a capability.
-In order to represent the unbounded address space, we use $-42$ to represent infinity.
-
-The $\jmp{}$ instruction updates the program counter to a requested location,
-but it is complicated by the presence of \emph{enter capabilities} 
-after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
-cannot be used to read, write or execute and their address and range cannot be
-modified. They can only be used to jump to. When that happens, their
-permission changes to $\exec$. They can be used to represent a kind of closures:
-an opaque package containing a piece of code together with local encapsulated
-state. Such a package can be built as an enter capability $c =
-((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
-contains local state (data or capabilities) and $[\addr,\addrend]$ contains
-instructions. The package is opaque to an adversary holding $c$. When $c$ is
-jumped to however, the instructions can start executing and have access to the local
-data through the updated version of $c$, now in the $\pcreg$-register. 
-\lau{As we talked about what is described here, is not how we make closures in
-the TR, so we should consider whether we should refrain from using that word
-here. }
-
-The instruction $\lea{}{}$ manipulates the current address of non enter-capabilities.
-It is fine for a capability to have a current address outside its range of authority as long as it is not used with an instruction that requires a specific capability.
-The instruction $\geta{}{}$ queries the current address of a capability and stores it in a register.
-
-Finally, the $\store{}{}$ instruction updates the memory to the argument value
-if the capability has write authority for the specified location. However, the
-instruction is complicated by the presence of \emph{local capabilities} modeled
-after the ones in the CHERI processor~\citep{Watson2015Cheri}. At a high-level, local
-capabilities are special in that they can only be kept in registers, i.e.\, they
-cannot be stored to memory. This means that local capabilities can be
-\emph{temporarily} given to an adversary, for the duration of an invocation. If
-we make sure to clear the capability from the register file after control is
-passed back to us, the adversary is unable to store the capability.
-However, there is one exception to the rule above: local capabilities can be
-stored to memory for which we have a capability with write-local authority
-(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack
-where register contents can be stored, including local capabilities. As long as
-all capabilities with write-local authority are themselves local and the stack
-is cleared after control is passed back by the adversary, we will see that this
-does not break the intended behavior of local capabilities.
-
-We point out that our local capabilities capture only a part of the semantics of
-local capabilities in CHERI. Specifically in addition to the above, CHERI's
-default implementation of the CCall exception handler forbids local capabilities
-from being passed across module boundaries. Such a restriction fundamentally
-breaks our calling convention since we pass around local return pointers and
-stack capabilities. However, CHERI's CCall is not implemented in hardware but
-in software precisely to allow experimenting with alternative models like ours.
-\dominique{Update to CHERI's new ccall story.}
-
-% Contrary to more traditional assembly languages, capability machines do not 
-% allow literal addresses to appear as part of instructions. This means that we
-
-In order to have a reasonably realistic system, we use a simple model of linking
-where a program has access to a linking table that contains capabilities for
-other programs. We also assume malloc to be part of the trusted computing base
-satisfying a certain specification. Malloc and linking tables are described
-further in the next section.
-\itoplas{The specification of malloc is presented in \sectionname~\ref{sec:malloc} as it uses the semantic model we build in Section~\ref{sec:logical-relation}.
-For full details on the linking table, we refer to the technical
-appendix~\citep{technical_appendix}.} \itoplassug{(Reviewer A, esop suggest explaining the
-  linking model)\\}
-\itoplassug{Suggestion: We also have the flag table used
-  for assertions, maybe we should add a description of it here.\\}
-% \itoplassug{Reviewer C, popl, would have liked a discussion of the relationship
-%   between this capability machines and old capability machines from the
-%   litterature (Cambridge CAP \& IBM System 39, see tex comment).\\}
- % * I'm old, so I'd appreciate a brief discussion LCM's relationship
- %  with the capability hardware I (and perhaps other older readers, or
- %  those outside this precise subfield) are likely to have heard of ---
- %  the Cambridge CAP & IBM System 39.  I think the capabilities of the
- %  machine are not that far from CAP, but this machine has the local
- %  capabilities support. Is that right?
-
-\section{Stack and Return Pointer Management Using Local Capabilities}
-\label{sec:stack-and-return-pointer}
-\itoplassug{Suggestion: say something about arguments in this paragraph instead of defering it to the discussion only?\\}
-\itoplassug{Suggestion: Add some of our presentation illustration figures to illustrate some of the issues? May not be feasible as they rely on ``animation''.\\}
-% ## Clarify attacker model earlier
-% The proposed calling convention seems able to protect multiple
-% components from each other, yet most of the paper discusses this
-% protection only in terms of one single statically known
-% "adversary". This seems to be selling short the work, and raising
-% questions of whether this calling convention is much weaker
-% security-wise than the CHERI one, which is clearly about protecting
-% multiple components. It would be good to explain this better upfront,
-% not just in the discussions section at the end.
-\itoplassug{Suggestion: Include a description of (possibly definition for) well-bracketedness?
-(requested by reviewer C, popl).
-Reviewer D, popl, asks for the same but also a definition of local state encapsulation.
-In the POPL notes, I wrote that I talked with Lars about it and this is something that is not easily defined.
-Perhaps we should have a discussion of this.\\}
-One of the contributions in this paper is a
-demonstration that local capabilities on a capability machine support a calling
-convention that enforces control-flow correctness in a way that is provably
-watertight, potentially efficient, does not rely on a trusted central stack
-manager, and supports higher-order interfaces to an adversary, where an adversary
-is just some unknown piece of code. In this section, we explain the high-level
-approach of this calling convention. We motivate each security measures with a
-situation we want to avoid (motivating each measure separately with a summary table at the end).
-After that, we define a number of reusable macro-instructions that can be used
-to conveniently apply the proposed convention in subsequent examples.
-
-The basic idea of our approach is simple: we stick to a single, rather standard,
-C stack and register-passed stack and return pointers much like a standard C
-calling convention. However, to prevent various ways of misusing this basic
-scheme, we put local capabilities to work and take a number of
-not-always-obvious safety measures. The safety measures are presented in terms
-of what \emph{we} need to do to protect ourselves against an \emph{adversary},
-but this is only for presentation purposes as our code assumes no special status on
-the machine. In fact, an adversary can apply the same safety measures to protect
-themselves against us. In the following paragraphs, we will explain the issues to be
-considered in all the relevant situations: when (1) starting our program, (2)
-returning to the adversary, (3) invoking the adversary, (4) returning from the
-adversary, (5) invoking an adversary callback, and (6) having a callback invoked
-by the adversary.
-
-\textbf{Program start-up} We assume that the language runtime initializes the
-memory as follows: a contiguous array of memory is reserved for the stack, for
-which we receive a stack pointer in the register $r_\stk$. We stress that
-the stack is not built-in, but merely an abstraction we put on this piece of the
-memory. The stack pointer is local and has $\rwlx$ permission. Note that this
-means that we will be placing and executing instructions on the stack.
-Crucially, the stack is the only part of memory for which the runtime (including
-malloc, loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
-Additionally, our examples typically also assume some memory to store
-instructions or static data. Another part of memory (called the heap) is
-initially governed by malloc and at program start-up, no other code has
-capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
-regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
-assume that memory allocated through malloc cannot be freed.\footnote{In more realistic settings, reusing freed memory on a capability machine can be made safe by checking or enforcing that there are no dangling pointers to the freed memory, as implemented, for example, in CHERI-JNI~\citep{chisnall_cheri_2017}.
-}
-
-\textbf{Returning to the adversary} Perhaps the simplest situation is returning
-to the adversary after they invoked our code \itoplas{(\figurename~\ref{fig:illu-intro})}. In this case, we have received a
-return pointer from them, and we just need to jump to it as usual. An obvious
-security measure to take care of is properly clearing the non-return-value
-registers before we jump (since they may contain data or capabilities that the
-adversary should not get access to). Additionally, we may have used the stack
-for various purposes (register spilling, storing local state when invoking other
-functions etc.), so we also need to clear that data before returning to the
-adversary \itoplas{(\figurename~\ref{fig:illu-priv-data} and \figurename~\ref{fig:illu-clear-priv})}.
-
-However, if we are returning from a function that has itself invoked adversary
-code \itoplas{(\figurename~\ref{fig:illu-call-adv})}, then clearing the used part of the stack is not enough. The \emph{unused}
-part of the stack may also contain data and capabilities, left there by the
-adversary, including local capabilities since the stack is write-local. As we
-will see later, we rely on the fact that the adversary cannot keep hold of local
-capabilities when they pass control to the trusted code and receive control
-back. In this case, the adversary could use the unused part of the stack to
-store local pointers and load them from there after they get control back
-\itoplas{(\figurename~\ref{fig:illu-adv-data})}. To
-prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
-the stack that the adversary has had access to; not just the parts that we have
-used ourselves \itoplas{(\figurename~\ref{fig:illu-clear-all})}. Since we may be talking about a large part of memory, this
-requirement is the most problematic aspect of our calling convention for
-performance (see the discussion in \sectionname~\ref{sec:discussion} for how this might be
-mitigated).
-
-\begin{figure}[ptb]
-  \centering
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-
-    % Unused part
-      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,5) rectangle (7.6,11.5);
-        \capbrace{(4.6,5)}{(4.6,12)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-    \caption{Just before we return to the adversary. Our stack frame
-      is the topmost, and we do not have access to the adversary stack frame.}
-  \label{fig:illu-intro}
-  \end{subfigure}
-\begin{subfigure}{0.033\linewidth}
-  % padding
-  \phantom{a}
-\end{subfigure}
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=red] {\footnotesize ostk};
-      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=red] {\footnotesize opc};
-      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=red] {\footnotesize ...};
-      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=red] {\footnotesize ...};
-
-    % Unused part
-      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace{(4.6,2)}{(4.6,12)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-    \caption{Just after we returned to the adversary but did not take
-      care to clear our local stack frame.}
-  \label{fig:illu-priv-data}
-  \end{subfigure}
-\begin{subfigure}{0.033\linewidth}
-  % padding
-  \phantom{a}
-\end{subfigure}
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5,color=teal] {\footnotesize 0};
-      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5,color=teal] {\footnotesize 0};
-      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5,color=teal] {\footnotesize 0};
-      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5,color=teal] {\footnotesize 0};
-
-    % Unused part
-      \node[opacity=0.5,rotate=-45] at (2.25,9) {\footnotesize Unused stack};
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace{(4.6,2)}{(4.6,12)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-    \caption{Just after we returned to the adversary, and we did take
-      care to clear our local stack frame.}
-  \label{fig:illu-clear-priv}
-  \end{subfigure}
-  \begin{subfigure}[t]{0.33\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \inactsf{(0,5)}{(4.5,7)}{Our stack frame}
-
-    % second attacker
-      \actadv{(0,7)}{(4.5,10)}{Callee\\\footnotesize stack frame}
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,7) rectangle (7.6,11.5);
-        \capbrace{(4.6,7)}{(4.6,13)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-  \caption{After an adversary has called us, and we have called the adversary.}
-  \label{fig:illu-call-adv}
-\end{subfigure}
-\begin{subfigure}[t]{0.02\linewidth}
-  % padding
-  \phantom{a}
-\end{subfigure}
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \draw[fill=white] (0,5) rectangle (4.5,5.5) node[pos=.5] {\footnotesize 0};
-      \draw[fill=white] (0,5.5) rectangle (4.5,6) node[pos=.5] {\footnotesize 0};
-      \draw[fill=white] (0,6) rectangle (4.5,6.5) node[pos=.5] {\footnotesize 0};
-      \draw[fill=white] (0,6.5) rectangle (4.5,7) node[pos=.5] {\footnotesize 0};
-
-    % Unknown part
-      \node[color=red] at (2.25,9) {\huge ?};
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace{(4.6,2)}{(4.6,13)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-  \caption{After the adversary has returned to us, and we have
-    returned from the first adversary call. At this point, the adversary has
-    access to the local state from the nested adversary call.}
-  \label{fig:illu-adv-data}
-\end{subfigure}
-\begin{subfigure}[t]{0.02\linewidth}
-  % padding
-  \phantom{a}
-\end{subfigure}
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % trusted stack frame
-      \actsf{(0,5)}{(4.5,7)}{Our stack frame}
-      \foreach \x in {5,5.5,...,10}
-      {
-        \draw[fill=white] (0,\x) rectangle (4.5,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
-      };
-
-      % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbrace{(4.6,2)}{(4.6,13)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-  \caption{After the adversary has returned to us, and we have
-    returned from the first adversary call after we took care to clear the
-    unused part of the stack.}
-  \label{fig:illu-clear-all}
-  \end{subfigure}
-  \caption{Depictions of the stack in relation to stack clearing.
-    The greyed out areas are parts of the stack that are not accessible at the
-    time.
-    The sp capability is the stack pointer.}
-  \label{fig:ret-adv2}
-\end{figure}
-\textbf{Invoking the adversary} A slightly more complex case is invoking the
-adversary. As above, we clear all the non-argument registers, as well as the
-part of the stack that we are not using (because, as above, it may contain local
-capabilities from previously executed code that the adversary could exploit in
-the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
-we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
-that we are not using ourselves.
-
-In one of the registers, we also provide a return pointer which must be a local
-capability. If it were global, the adversary would be able to store away the
-return pointer in a global data structure (i.e.\ there exists a global capability for it) and jump to it later in
-circumstances where this should not be possible. For example, they could store
-the return pointer, legally jump to it a first time, wait to be invoked again,
-and then jump to the old return pointer a second time instead of the new return
-pointer received for the second invocation. Similarly, they could store the
-return pointer, invoke a function in our code, wait for us to invoke them again,
-and then jump to the old return pointer rather than the new one received for the
-second invocation.
-% \lau{If we need space, then maybe one of the examples is sufficient.}
-% \lau{Both of these are good to have their as the two last examples of \sectionname~\ref{sec:examples} use them.}
-%
-% \lau{In the examples, we refer to the local return pointer as
-%   ``protected return pointer''. We should either introduce that here
-%   or use a different formulation in the examples. }
-% \lau{It seemed did not seem to add anything to adopt the terminlogy - we
-%   often highlight that the return pointer is local which in most cases
-%   is so important that we do not want to leave it out.}
-By making the return pointer local, we prevent such attacks.
-The adversary can only store local capabilities with a write-local capability, and the only piece of memory governed by a write-local capability is the stack.
-Since the stack pointer itself is also local, it can also only be stored on the stack.
-There is no way for the adversary to recover either of these local capabilities because we clear the part of the stack that they had access to before we pass control back to them.
-
-Note that storing stack pointers for use during future invocations would also be
-dangerous in itself, i.e. not just because it can be used to store return
-pointers. Imagine that the adversary stores their stack pointer
-(\figurename~\ref{fig:illu-adv-store-sp}), invokes trusted code
-that uses part of the stack to store private data
-(\figurename~\ref{fig:illu-adv-call}) and then invokes the adversary
-again with a stack pointer restricted to exclude the part containing the private
-data (\figurename~\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
-could access the private data stored there by the trusted code and break
-local-state encapsulation.  
-% \lau{Do we say anything about arguments to the callback?
-Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary.
-Can we even to any extend allow local arguments that we pass on from an unknown source?}
-% Domi: this is now discussed in the discussion section, no need to discuss it here?
-
-\begin{figure}[htb]
-  \centering
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \actadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
-    % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbracebot{(4.6,2)}{(4.6,12)}{sp}
-      \end{scope}
-    \end{tikzpicture}
-    \caption{The stack after the adversary has stored its stack pointer
-      far up on the stack.}
-  \label{fig:illu-adv-store-sp}
-\end{subfigure}
-\begin{subfigure}{0.043\linewidth}
-  % padding
-  \phantom{a}
-\end{subfigure}
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
-
-    % Our stack frame
-      \actsf{(0,5)}{(4.5,6.5)}{Our stack frame}
-
-    % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbracebotred{(4.6,2)}{(4.6,12)}{sp}
-      \end{scope}
-      \begin{scope}
-        \clip (5.1,5) rectangle (9,11.5);
-        \capbrace{(5.1,5)}{(5.1,12)}{sp'}
-      \end{scope}
-
-    \end{tikzpicture}
-  \caption{The stack after the adversary has called some trusted code.}
-  \label{fig:illu-adv-call}
-  \end{subfigure}
-\begin{subfigure}{0.043\linewidth}
-  % padding
-  \phantom{a}
-\end{subfigure}
-  \begin{subfigure}[t]{0.3\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-    \stdstackstart
-
-    % attacker ..
-      \inactadv{(0,2)}{(4.5,5)}{Adv. stack frame}
-
-    % Stored stack pointer
-      \draw[fill=white] (0,10) rectangle (4.5,10.5) node[pos=.5,color=red] {\footnotesize sp};
-
-    % Our stack frame
-      \inactsf{(0,5)}{(4.5,6.5)}{Our stack frame}
-
-    % Second attacker
-      \actadv{(0,6.5)}{(4.5,9.5)}{Callee \\ \footnotesize stack frame}
-
-    % Stack pointer
-      \begin{scope}
-        \clip (4.6,2) rectangle (7.6,11.5);
-        \capbracebotred{(4.6,2)}{(4.6,12)}{sp}
-      \end{scope}
-      \begin{scope}
-        \clip (5.1,6.6) rectangle (9,11.5);
-        \capbrace{(5.1,6.5)}{(5.1,12)}{sp'}
-      \end{scope}
-    \end{tikzpicture}
-  \caption{The stack after the trusted code has called the adversary. The
-    adversary's old stack pointer is still available on the stack.}
-  \label{fig:illu-adv-called}
-  \end{subfigure}
-  \caption{Illustration of the situations related to stack clearing when invoking an adversary.
-    The greyed out areas are parts of the stack that are not accessible at the time.
-  }
-  \label{fig:invoke-adv}
-\end{figure}
-
-\textbf{Returning from the adversary} Return pointers must be passed as local
-capabilities. But what should their permissions be, what memory should they
-point to and what should that memory (the activation record) contain? Let us
-answer the last question first by considering what should happen when the
-adversary jumps to a return pointer. In that case, the program counter should be
-restored to the instruction after the jump to the adversary, so the activation
-record should store this old program counter. Additionally, the stack pointer
-should also be restored to its original value. Since the adversary has a more
-restricted authority over the stack than the code making the call, we cannot
-hope to reconstruct the original stack pointer from the stack pointer owned by the
-adversary. Instead, it should be stored as part of the activation record.
-
-\begin{figure}[b]
-  \begin{tikzpicture}[scale=.7, every node={scale=.7}]
-    \draw[fill=gray!20,draw=none] (0,4) rectangle (6,3.5);
-    \draw[fill=gray!20,draw=none] (0,0.5) rectangle (6,1);
-    \draw[draw=gray!80] (0,0.5) -- (0,4);
-    \draw[draw=gray!80] (6,0.5) -- (6,4);
-    \draw (0,1) rectangle (6,1.5) node[pos=.5] {\footnotesize previous stack pointer};
-    \draw (0,1.5) rectangle (6,2) node[pos=.5] {\footnotesize previous program counter};
-    \draw (0,2) rectangle (6,3.5) node[pos=.5] {\footnotesize restore instructions};
-    \draw (-3,3.25) node (rp) {return pointer};
-    \draw[->] (rp.east) -- node[above] {$\entry$}(-.2,3.25);
-  \end{tikzpicture}
-
-  \caption{Structure of an activation record}
-  \label{fig:activ-rec-struct}
-\end{figure}
-
-Clearly, neither the old program counter nor the old stack pointer should be accessible by the adversary. In
-other words, the return pointer provided to the adversary must be a capability
-that they can jump to but not read from, i.e.\ an enter capability. To make this
-work, we construct the activation record as depicted in
-\figurename~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
-over the entire activation record (containing the previous return and stack
-pointer) and its current address points to a number of restore instructions in
-the record. Upon invocation, the instructions in the activation record are executed and can
-load the old stack pointer and program counter back into the register file. As
-the return pointer is an enter pointer, the adversary cannot get hold of the
-activation record's contents. However after invocation, its permission is updated to
-$\exec$, so the contents become available to the restore instructions.
-
-The final question that remains is: where should we store this activation
-record? The attentive reader may already see that there is only one possibility:
-since the activation record contains the old stack pointer, which is local, the
-activation record can only be constructed in a part of memory where we have
-write-local access, i.e.\ on the stack. Note that this means we will be placing
-and executing instructions on the stack, i.e.\ it will not just contain code
-pointers and data. This means that our calling convention should be combined
-with protection against stack smashing attacks (i.e.\ buffer overflows on the
-stack overwriting activation records' contents). Luckily, the capability
-machine's fine-grained memory protection \citep{Woodruff:2014:CCM:2665671.2665740} should make it reasonably easy for a
-compiler to implement such protection by making sure that only appropriately
-bounded versions of the stack pointer are made available to source language
-code.
-
-\textbf{Invoking an adversary callback} If we have a higher-order interface to
-the adversary, we may need to invoke an adversary callback. In this case, not so
-much changes with respect to the situation where we invoke static adversary
-code. The adversary can provide a callback as a capability for us to jump to,
-either an $\entry$-capability if they want to protect themselves from us or just
-an $\exec$ capability if they are not worried about that. However, there is one
-scenario that we need to prevent: if they construct the callback capability to
-point into the stack, it may contain local capabilities that they should not
-have access to upon invocation of the callback. As before, this includes return
-and stack pointers from previous stack frames that they may be
-trying to illegally use inside the callback.
-
-To prevent this, we only accept callbacks from the adversary in the form of
-global capabilities, which we dynamically check before invoking them (and we
-fail otherwise). This should not be an overly strict requirement: our own
-callbacks do not contain local data themselves, so there should be no need for
-the adversary to construct callbacks on the stack.\footnote{Note that it does
-  prevent a legitimate but non-essential scenario where the adversary wants to
-  give us temporary access to a callback not allocated on the stack.}
-\itoplassug{Maybe a word on arguments passed from the adversary to us that we pass on to a callback given by the adversary (mentioned by reviewer E, popl, see tex comment.)}
-%page 8: "we only accept callbacks from the adversary in the form of global capabilities, which we dynamically check before invoking them".
-This makes sense, but don't you also need to check arguments that are passed from the adversary to us and then passed by us into the callback?
-
-\textbf{Having a callback invoked by the adversary} The above leaves us with
-perhaps the hardest scenario: how to provide a callback to the adversary. The
-basic idea is that we allocate a block of memory using malloc that we fill with
-the capabilities and data that the callback needs, as well as some
-prelude instructions that load the data into registers and jumps to the right
-code. Note that this implies that no local capabilities can be stored as part of
-a closure. We can then provide the adversary with an enter-capability covering
-the allocated block and pointing to the contained prelude instructions. However,
-the question that remains in this setup is: from where do we get a stack pointer when
-the callback is invoked?
-
-Our answer is that the adversary should provide it to us; just as we provide
-them with a stack pointer when we invoke their code. However, it is important
-that we do not just accept any capability as a stack pointer but check that it
-is safe to use. Specifically, we check that it is indeed an $\rwlx$ capability. Without
-this check, an adversary could potentially get control over our local stack
-frame during a subsequent callback by passing us a local $\rwx$ capability to a global data
-structure instead of a proper stack pointer
-and a global callback for our callback to invoke. If our local state contains no
-local capabilities, then, otherwise following our calling convention, the callback would
-not fail and the adversary could use a stored capability for the global data
-structure to access our local state. To prevent this from happening, we need to
-make sure the stack capability carries $\rwlx$ authority since the system wide assumption then
-tells us that the adversary cannot have global capabilities to our local stack.
-
-\textbf{Calling convention} With the security measures introduced and motivated,
-let us summarize our proposed calling convention:
-\begin{description}[font=\normalfont\itshape]
-\item[At program start-up] A
-local $\rwlx$ stack pointer resides in register $r_\stk$. No global write-local capabilities.
-\item[Before returning to the adversary] Clear non-return-value
-registers. Clear the part of the stack we had access to (not just the part we used).
-\item[Before invoking the adversary] Push activation record to the
-stack. Create return pointer as local $\entry$-capability to the instructions in
-the record.  Restrict the stack capability to the unused part and clear
-it. Clear non-argument registers.
-\item[Before invoking an adversary callback] Make sure callback is global.
-  \item[When invoked by an adversary] Make sure received stack pointer has permission $\rwlx$.
-\end{description}
-
-%\itoplassug{Suggestion: Make it more clear that this CC can be used to protect multiple components (reviewer A, ESOP), see tex comment.\\}
-\begin{toplas}
-  \textbf{Modularity} The calling convention ensures well-bracketed calls and
-  local-state encapsulation for the caller but not the callee. In the above
-  presentation to make it easy to distinguish between the parties involved, we
-  present the callee as some adversarial code that we do not trust. In reality,
-  the callee could be well-behaved and wish to ensure well-bracketed calls and
-  local-state encapsulation as well. The calling convention puts no restriction
-  on the callee that the caller itself does not follow, so by following the
-  calling convention, the callee can also obtain those guarantees. In other
-  words, the calling convention is modular and scales to scenarios with multiple
-  distrusting parties invoking each other.
-\end{toplas}
-
-\section{Logical Relation}
-\label{sec:logical-relation}
-\itoplassug{Suggestions for things to add heres
-  \begin{itemize}
-  \item (The lemma is there, consider pointer) Double monotonicity lemma (lemma 77 and 79) with a pointer back to local capability section of popl intro to this section.
-  \item Reviewer B, ESOP suggests inferring general security properties from the LR.
-We had a section in the discussion that explained why this was not really possible.
-I have reintroduced it in the discussion.
-  \end{itemize}
-}
-
-% In this section, we formalize the guarantees provided by the
-% capability machine, including the specific guarantees for local
-% capabilities, by means of a step-indexed Kripke logical relation 
-% with recursively defined worlds. 
-% We use the logical relation in the following section to show
-% local-state encapsulation and control-flow integrity properties for
-% challenging example programs.
-
-\begin{toplas}
-%\todo[inline]{Look this introduction and the following section through to check for overlap.
-Also consider making backwards references to this introduction from the places where we use advanced techniques, so it becomes clear that they are necessary.}
-% Our logical relation will prescribe which values are capability safe given a
-% world $W$ containing a set of memory invariants. For expressiveness,
-% we use invariants which can evolve over time (sometimes also referred
-% to as \emph{protocols}). We thus start by describing the worlds we
-% will use.
-Now that we have defined our calling convention, how can we be sure that it
-works? More concretely, suppose that we have a program that uses the convention
-in its interaction with untrusted adversary code. Can we formally prove the
-program's correctness if it relies on well-bracketed control flow and private
-state encapsulation for the interaction with the adversary? Clearly, such a
-proof should depend on a formal expression of the guarantees provided by the
-capability machine, including the specific guarantees for local capabilities.
-
-In this section, we construct such a formalization. We make use of some
-well-studied and powerful (but non-trivial) machinery from the literature.
-Specifically, we employ a unary step-indexed Kripke logical relation with
-recursive worlds and some additional special characteristics of our own.
-Step-indexing, Kripke logical relations, and recursive worlds are techniques that
-may be familiar from lambda calculus settings, but it may not be clear to the
-reader how they apply in this more low-level assembly language. Therefore, in the next section, we do
-not immediately dive into the details, but first we try to provide some informal
-intuition about how all of this machinery comes into play in our setting.
-
-Note: even though the calling convention is the main application in this paper,
-the logical relation we construct is very general and can be seen as a
-formulation of capability safety; hence it should be regarded as an
-independent contribution.
-
-\subsection{Formalizing the guarantees of the capability machine}
-\label{sec:formalizing-guarantees}
-What differentiates a capability machine from a more standard assembly language
-is that we can bound the authority of an executing block of code based solely
-on the capabilities it has access to. Specifically, it does not matter which
-instructions are actually executed, i.e.\ the bound also applies to untrusted
-adversary code that has not been inspected or modified in any way.
-
-\paragraph{Worlds}
-But what does a ``bound on the authority'' of an executing block of code mean?
-In our setting, there are no externally observable side-effects; the only
-primitive authority that code may hold is authority over memory. As such, the
-authority bounds we consider are related to memory, but in a form that is more
-fine-grained than standard read/write authority: a piece of code's authority can be bounded
-by arbitrary memory invariants that it is required to respect. Specifically, we
-will define worlds $W \in \Worlds$, which describe a set of memory invariants, and
-our results will express authority bounds on code as \emph{safety with respect to such
-a world,} i.e.\ the fact that the code respects the invariants registered in
-the world.
-
-\paragraph{Safe values}
-Suppose we have a world $W$ expressing that the memory must contain
-value $42$ at address $0$, may contain arbitrary values at
-addresses $50$-$60$, a $\readwrite$ capability for address $0$ at address $73$, and
-an integer at address $100$ that may only increase over time\footnote{Indeed, we
-  will allow a notion of \emph{evolvable} invariants, aka \emph{protocols}, that can
-  express such a temporal property.}. Our main theorem will state that if the
-current register file only contains safe words (numbers or capabilities which
-preserve the invariants in $W$ under any interaction), then an execution will
-necessarily also preserve the memory invariants (irrespective of the instructions being
-executed).
-
-To make this more precise, we need to define the set $\stdvr(W) \in
-\powerset{\Words}$ of words that are safe w.r.t.\ $W$. Essentially, the set
-should only include words that preserve $W$'s invariants under any interaction,
-but should otherwise be as liberal as possible. Numbers are clearly always safe
-as they cannot be used to break invariants. Whether a capability is safe depends
-on the authority that it carries. In the above-described world, a read capability
-for address $0$ is safe as it can only be used to read the value $42$, which is
-itself safe. However, a write capability for address $0$ is not safe: it can be
-used to overwrite the memory at that address with a value other than $42$
-breaking the invariant for that address.
-
-\paragraph{Step-indexing}
-More generally, we want to define that a read capability for memory range
-$[b,e]$ is safe if the world guarantees that the words at those addresses are
-themselves safe. However, this definition is cyclic: suppose the world
-guarantees that the memory at address $a$ will contain a read capability for
-address $a$? Then the definition says that a read capability for address
-$a$ is safe if and only if the same read capability for address $a$ is safe. This form of
-cyclic reasoning is related to similar challenges in languages with recursive
-types or higher-order ML-style references, and a standard solution is
-to use step-indexing \citep{Appel:2001:IMR:504709.504712}: essentially, the cycle is broken by defining safety up to a
-certain number of interaction steps. All words will be considered
-safe up to $0$ steps (since if there is no interaction, nothing unsafe
-can happen), and, for example, a read capability will be safe up to $n$
-steps if the world guarantees that the words at the corresponding addresses are
-safe up to $n - 1$ steps. We can then prove that the above read capability for
-address $a$ is safe up to any number of steps.
-
-\paragraph{Future worlds}
-Worlds are defined as a set of invariants on the memory, but what if we
-allocate fresh memory through malloc? We may want to establish new invariants
-for this freshly allocated memory and be sure that the adversary will also
-respect those (if we don't provide them with capabilities through which the new
-invariants can be broken). To accommodate this, we allow worlds to evolve, for
-example by adding additional invariants for freshly allocated memory. Formally,
-we define valid ways for a world $W$ to evolve into a new world $W'$ through a
-future-world relation $W' \future W$ and we ensure that the set of safe words in
-world $W$ must remain safe in any future world $W'$. Defining safety w.r.t.\ a
-notion of evolvable worlds makes our logical relation into a \emph{Kripke}
-logical relation \citep{pitts_operational_1998}.
-
-\paragraph{Invariants and Recursive Worlds}
-Worlds group a set of memory invariants, but how are they actually
-defined formally? We represent each invariant by a region
-$r \in \Regions$. We will see later that regions contain state
-machines to support a notion of evolvable invariants. Every
-state of the state machine contains a predicate $H$ that defines the valid
-memory segments.
-Unfortunately, it is not enough to just take
-$H \in \powerset{\MemSegments}$, because sometimes the invariant may
-itself be world-dependent. For example, we may want to express
-invariants like ``the memory at address $50$ contains a value that is
-safe in the current world''. As explained, worlds may evolve, and the
-set of safe values may grow in future worlds. Therefore, we need to
-index $H$ over worlds, i.e., take
-$H \in \Worlds \rightarrow \powerset{\MemSegments}$. The result is worlds containing regions with world-indexed predicates,
-i.e., the set of worlds must be recursively defined. We will see how such a recursive
-definition can be accommodated using techniques from the
-literature (essentially an advanced application of step-indexing).
-
-\paragraph{Local capabilities}
-When we invoke an untrusted piece of code and provide it with certain global capabilities, it may have stored those capabilities in memory.
-In this case, we will only be able to reinvoke the code if we can guarantee that those values are still valid.
-Formally, worlds represent the invariants that global capabilities' safety relies on and the reinvocation is only safe in future worlds where the invariants are respected.
-
-However, if we provide the adversary with local capabilities in that first invocation, then the situation is a bit different.
-The adversary has no way to store these local capabilities, so if we make sure that there are also no old local capabilities in the register file for the second invocation (including the capability being invoked), then the adversary cannot use them any more.%
-\footnote{We ignore write-local capabilities in this discussion.
-  If the adversary does have access to write-local capabilities in the first and second invocation, then the memory they address must be cleared entirely before the second invocation in order for the reinvocation to remain safe.}
-Therefore, we can allow the second invocation to happen in any \emph{private} future worlds ($W' \futurestr W$) in which safe global capabilities remain safe but local capabilities do not.
-This private future world relation is more liberal than the standard public one ($W' \futurewk W$, in which \emph{all} safe capabilities remain safe).
-Concretely, worlds may contain \emph{temporary} regions representing invariants that only local capabilities may rely on for their safety and that may be revoked (disabled) in private future worlds.
-
-Interestingly, this idea is a variant of a notion of public/private future worlds that has been previously used in the literature (see \sectionname~\ref{sec:discussion}).
-However, temporary regions are new in our setting and there is an interesting interplay with the recursiveness of the worlds: for a temporary region, the predicate $H \in \Worlds \rightarrow \powerset{\MemSegments}$ (which defines the safe memory segments in the current world) is only required to be monotone w.r.t.\ public future worlds (i.e.\ safe memory segments remain safe in public future worlds).
-On the other hand for permanent regions, the world-indexed predicate must be monotone w.r.t.\ private future worlds.
-As a consequence, the memory for a permanent region may not contain local capabilities (as their safety would be broken in private future worlds), which in turn implies that only local capabilities may have write-local permission (a general sanity requirement when using local capabilities\footnote{In fact, local capabilities become useless as soon as the adversary has access to a single global, write-local capability.})\footnote{For explanation purposes, this discussion ignores certain ways to allow for local capabilities in a permanent region, for example, by not requiring that they are valid or requiring that they are local versions of valid global capabilities.}.
-\end{toplas}
-\subsection{Worlds}
-\label{subsec:worlds}
-\begin{figure}[bt]
-  \centering
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    \scope
-    \clip (-.1,-.1) rectangle (4.6,11.1);
-    \fill[fill=white] (0,0) rectangle (4.5,11);
-    \draw (0,0) -- (0,11);
-    \draw (4.5,0) -- (4.5,11);
-    \endscope
-    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-    \draw (-1.5,0) node {};
-    \draw (-1.5,15) node {};
-
-    \foreach \x in {0.5,1,...,4.5}
-    {
-      \shade[left color=blue!70,right color=white] (0,\x) rectangle (4.5,\x+.5);
-      \draw (0,\x) rectangle (4.5,\x+.5);
-    };
-    \foreach \x in {5,5.5,...,7}
-    {
-      \fill[color=red!55] (0,\x) rectangle (4.5,\x+.5);
-      \draw (0,\x) rectangle (4.5,\x+.5);
-    };
-    \foreach \x in {7.5,8,...,10}
-    {
-      \fill[color=green!85] (0,\x) rectangle (4.5,\x+.5);
-      \draw (0,\x) rectangle (4.5,\x+.5);
-    };
-
-    \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=0pt},yshift=0pt]
-    (4.6,1.5) -- (4.6,3) node (a) [black,midway] {};
-    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.3},yshift=0pt]
-    (-0.1,0.5) -- (-0.1,3) node (b) [black,midway,xshift=0.8cm] {};
-    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt},yshift=0pt]
-    (-0.1,3) -- (-0.1,4.5) node (c) [black,midway,xshift=0.8cm] {};
-    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt},yshift=0pt]
-    (-0.1,5) -- (-0.1,7) node (d) [black,midway,xshift=0.8cm] {};
-    \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.6},yshift=0pt]
-    (-0.1,7.5) -- (-0.1,10.5) node (e) [black,midway,xshift=0.8cm] {};
-
-    \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
-    {\fill[black] circle(1.5pt);}}}}
-
-    \draw[capa] (-4,8.5) edge[-latex,postaction={decorate},in=180,out=30,densely dashed] (-0.4,9.3);
-
-    \draw[capa] (-4,6.8) edge[-latex,postaction={decorate},in=180,out=-30] (-0.4,6);
-    \draw[capa] (-4,4.55) edge[-latex,postaction={decorate},in=180,out=-30] (-0.4,3.75);
-
-    \draw[capa] (-4,2.05) edge[-latex,postaction={decorate},in=180,out=-30,densely dashed] (-0.4,1.25);
-
-    \draw[capa] (4.25,6.25) edge[-latex,postaction={decorate},in=0,out=0] (4.9,2.25);
-
-    \draw (-2,4.25) edge[color=red,thick] (-1,3.25)
-          (-1,4.25) edge[color=red,thick] (-2,3.25);
-    \draw (5.35,3.85) edge[color=red,thick] (6.35,2.85)
-          (6.35,3.85) edge[color=red,thick] (5.35,2.85);
-
-    \node at (-8,10) (perm) {\footnotesize Permanent region};
-    \node at (-8,9) (temp) {\footnotesize Temporary region};
-    \node at (-8,8) (temp) {\footnotesize Local capability};
-    \node at (-8,7) (temp) {\footnotesize Global capability};
-
-    \fill[color=green!85] (-12.5,9.75) rectangle (-11.5,10.25);
-    \fill[color=red!55] (-11.5,9.75) rectangle (-10.5,10.25);
-    \fill[color=green!85] (-11.75,9.75) -- (-11.75,10.25) -- (-11.25,10.25);    
-    \fill[color=red!55] (-11.75,9.75) -- (-11.25,9.75) -- (-11.25,10.25);
-    \draw (-12.5,9.75) rectangle (-10.5,10.25);
-
-    \shade[left color=blue!70,right color=white] (-12.5,8.75) rectangle (-11.5,9.25);
-    \draw (-12.5,8.75) rectangle (-10.5,9.25);
-
-    \draw[capa] (-12.4,8) edge[-latex,postaction={decorate},in=180,out=30,densely dashed] (-10.5,8);
-    \draw[capa] (-12.4,7) edge[-latex,postaction={decorate},in=180,out=30] (-10.5,7);
-    \draw (-13,10.6) rectangle (-5.4,6.4);
-  \end{tikzpicture}
-
-  \caption{The relation between local/global capabilities and temporary/permanent
-    regions. The colored fields are regions governing parts of memory. Global capabilities cannot depend on temporary
-    regions.}
-  \label{fig:cap-world}
-\end{figure}
-% \dominique{Perhaps we could first try to convey the intuition of a permanent
-%   world with a temporary overlay, with global capabilities ignoring the overlay.
-%   Perhaps we could add a depiction of the world to Figure 5?}
-%\todo[inline]{Find a good place to describe monotonicity of
-%interpretation function.}
-% \lau{this may not be an important enough point to include it.}
-
-%As describe above, worlds are a collection of invariants. Formally, 
-A world is a semantic model of the memory that carves out memories with a particular shape from all the possible memories.
-In our correctness proofs, worlds allow us to rely on the memory having a particular shape, but sometimes we will also have to guarantee that the memory has a certain shape.
-Essentially, a world is a collection of invariants.
-The memory satisfies the world when part of the memory satisfy each of the invariants.
-
-Worlds are represented as a finite map from region names, modeled as natural numbers, to regions that each correspond to an invariant on part of the memory.
-We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
-Each permanent and temporary region contains a state transition system with public and private transitions that describe how the invariant is allowed to change over time.
-In other words, they are protocols for the region's memory.
-Protocols imposed by permanent regions stay in place indefinitely.
-Any capability, local or global, can depend on these protocols.
-Protocols imposed by temporary regions can be revoked in private future worlds.
-Doing this may break the safety of local capabilities but not global ones.
-This means that local capabilities can safely depend on the protocols imposed by temporary regions, but global capabilities cannot since a global capability may outlive a temporary region that is revoked.
-This is illustrated in \figurename~\ref{fig:cap-world}.
-
-We need the future world relation to be extensional, so we do not actually remove a revoked temporary region from the world, but we turn it into a special revoked region that exists for this purpose.
-Such a revoked region contains no state transition system and puts no requirements on the memory.
-It simply serves as a mask for a revoked temporary region.
-Masking a region like this goes back to earlier work of \citet{Ahmed2004semantics} and was also used by \citet{Thamsborg:2011:KLR:2034773.2034831}.
-
-% Recursive worlds intuition
-% The worlds end up being defined recursively: The memory can contain
-% capabilities,
-% which grant authority over parts of memory. A world
-% models memory, so it also needs to be able to model capabilities in
-% memory and the protocols on them. Whether a capability satisfies the
-% protocol of a region may depend on what it grants authority over in
-% memory which is described by the world.
-
-%\dominique{``end up being defined recursively'' does not sound very insightful}
-% A region defines what its part of memory can legally look like. However,
-% validity of the memory's contents may itself depend on the state of other parts
-% of memory. To accommodate this, worlds need to be defined recursively: Worlds
-% describe memory, which can contain capabilities, whose safety is
-% world-dependent. 
-%As described in \sectionname~\ref{sec:formalizing-guarantees}, regions define sets of
-
-Regions are used to define safe memory segments, but this set may itself be world-dependent.
-In other words, our worlds are defined recursively.
-Recursive worlds are common in Kripke models and the following theorem uses the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for constructing them.
-The formulation of the lemma is technical, so we recommend that non-expert readers ignore the technicalities and accept that there exists a set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying the (recursive) equations in the theorem (where the $\blater$ operator can be safely ignored\footnote{The interested reader can find a brief coverage of \cofe{}'s and $\blater$ in Appendix~\ref{app:cofe}.}).
-% Short technical explanation
-% The recursiveness of the worlds gives rise to a recurisve domain
-% equation. Theorem~\ref{thm:world-existence} gives the solution to this
-% recursive domain equation.
-% \dominique{it is not explained what cofes are?}
-% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
-\begin{theorem}\label{thm:world-existence}
-  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
-  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
-  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
-  \begin{align*}
-      \xi \; : \; & \Wor \cong \blater (\nats \finparfun \Regions)\\
-      \Regions  = & \; \{\revoked\} \uplus \\
-               & \; \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))\uplus \\
-               & \; \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
-    \end{align*}
-  and for $W, W' \in \Wor\ldotp \quad  
-  \begin{aligned}
-    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
-    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
-  \end{aligned}$
-\end{theorem}
-In the above theorem, $\States \times \Rels$ corresponds to the aforementioned
-state transition system where $\Rels$ contains pairs of relations corresponding
-to the public and private transitions, and $\States$ is a set of world states that
-we assume to at least contain the states we use in this paper. The last part of
-the temporary and permanent regions is a state interpretation function that
-determines what memory segments the region permits in each state of the state
-transition system.  The different monotonicity requirements in the two
-interpretation functions reflect how permanent regions rely only on permanent
-protocols whereas temporary regions can rely on both temporary and permanent
-protocols.  $\UPred{\HeapSegments}$ is the set of step-indexed, downwards closed
-predicates on memory segments:
-$\UPred{\HeapSegments} = \{ A \subseteq  \nats \times
-\HeapSegments \mid 
-\forall (n, ms) \in A. \forall m
-\leq n\ldotp (m, ms) \in A\}$. 
-\
-
-With the recursive domain equation solved, we could take $\Wor$ as our notion of
-worlds, but it is technically more convenient to work with the following
-definition instead:
-\[
-  \Worlds = \nats \finparfun \Regions
-\]
-
-\subsubsection{Future Worlds}
-\label{subsec:future-worlds} 
-The future world relations model how memory may evolve over time.
-We have a \emph{public} and a \emph{private} future world relation that respectively what changes any capability can rely on and what changes only temporary regions can rely on.
-As temporary capabilities fall within the category of all capabilities, the public future relation is subsumed in the private future relation.
-
-The public future world $W' \futurewk W$ requires that $\dom(W') \supseteq \dom(W)$ and $\forall r \in \dom(W) \ldotp W'(r) \futurewk W(r)$.
-That is in a public future world, new regions may have been allocated, and existing regions may have evolved according to the public future region relation (defined below).
-The private future world relation $W' \futurestr W$ is defined similarly, using a private future region relation.
-The public future region relation is the simplest.
-It satisfies
-% \dominique{is defined by?}
-% \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations.
-It can be shown that these will have the properties mentioned below}
-the following properties:
-\begin{mathpar}
-  \inferrule{ (s,s') \in \phi_\pub }
-  {  (v,s',\phi_\pub,\phi,H) \futurewk (v,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
-  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
-  \and
-  \inferrule{ }
-  { \revoked \futurewk \revoked }
-\end{mathpar}
-In public future worlds, both temporary and permanent regions are only allowed to transition according to the public part of their transition system.
-Additionally, revoked regions must either remain revoked or be replaced by a temporary region.
-This means that the public future world relations allows us to reinstate a region that has been revoked earlier.
-The private future region relation satisfies:
-\begin{mathpar}
-  \inferrule{ (s,s') \in \phi} 
-  {  (v,s',\phi_\pub,\phi,H) \futurestr (v,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ r \in \Regions }
-  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
-  \and
-  \inferrule{ r \in \Regions }
-  { r \futurestr \revoked }
-\end{mathpar}
-Here, revocation of temporary regions is allowed.
-In fact, temporary regions can be replaced by an arbitrary region not just the special $\revoked$.
-Conversely, $\revoked$ regions may also be replaced by any other region.
-On the other hand, permanent regions cannot be masked away.
-They are only allowed to transition according to the private part of the transition system.
-
-\begin{toplas}
-  In order to be able to solve the recursive domain equation, we want the future world relations to be extensional.
-  Intuitively, the future world relation specifies how memory protocols may change over time.
-  If the future world relation was not extensional, then the protocols could just disappear entirely regardless of what they promise.
-  A piece of memory can still be repurposed under the right circumstances, i.e., if the protocol allows it under the future world relation used.
-  In practice, the extensionality of future worlds means that a future world cannot forget about region names in a past world.
-  This is the reason we use masks rather than deleting the regions altogether.
-  It does not, however, matter what kind of region is used as a mask; it is the region name we must retain.
-  This is why we allow revoked regions to be replaced by other regions in the two future world relations.
-  This may seem unnecessary as the future world relation allows new regions to be added.
-  However, for technical reasons in private future worlds, we need to be able to reinstate old regions with the same region names as they had in a past world.
-  Further, we do not want public future worlds to be able to prevent reinstation in private future worlds.
-  Private future worlds cannot revoke permanent regions, so we cannot allow public future worlds to use permanent regions as masks for revoked regions.
-\end{toplas}
-
-%domi: the below seems hard to understand at this moment.
-% The public future region relation only allows revoked regions
-% to be reused as temporary region. This means that it is always possible to turn
-% it back into a revoked region in a private future region. This requirement turns
-% out to be important in order to be able to restore the invariants of a caller.
-
-%%%%%%%%
-
-% In order to model changes to memory over time, we use the future world
-% relations. The future world relations model how more memory may be
-% allocated but also how memory may change. How memory is allowed to
-% change over time depends on the region that governs it and the future
-% world relation used. The most interesting part about the future worlds
-% is how it affects the regions, so we first look at the properties of
-% the \emph{public future region} relation and the \emph{private future
-%   region} relation. We then use these relations to describe the future
-% world relations.
-
-% {\bf Private future regions}
-% In the private future region relation, we allow revocation of $\temp$
-% regions. The regions cannot be revoked by removing them from the
-% world, because we need the future world relations to be
-% monotone. This is why we have the $\revoked$ region, so we have
-% a region that can serve as a mask. We can, however, let any region
-% serve as a mask, so in the private future region relation, we allow a
-% $\temp$ region to become any valid region. As the $\revoked$ region
-% serves as a mask, we let any valid region take its place in the
-% private future region relation.
-
-% If a $\temp$ region is not masked by a region, then it behaved like a
-% $\perm$ region. The $\perma$ regions stay the same type and they are
-% allowed to take a transition in the private part of the transition
-% system. Which can be thought of as the memory changes according to
-% some protocol.
-
-% The properties of the \emph{private future region} relation are:
-% \begin{mathpar}
-%   \inferrule{  (s,s') \in \phi \\
-%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-%   {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
-%   \and
-%   \inferrule{ r \in \Regions }
-%   { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
-%   \and
-%   \inferrule{ r \in \Regions }
-%   { r \futurestr \revoked }
-% \end{mathpar}
-
-% {\bf Public future regions}
-% In the public future regions we do not permit revocation of $\temp$
-% regions. In this relation, the $\temp$ and $\perma$ regions must stay
-% the same type and they are allowed to take a transition in the public
-% part of the transition system.
-
-% The $\revoked$ regions serve as masking regions, but are not as
-% liberal as in the private future region relation when it comes to
-% letting other regions replace them as masks. In the public future
-% region relation, we only allow $\revoked$ and $\temp$ regions to take
-% on the job as masking region. We have this limitation, because we
-% want the private future region relation to be able to reuse any region
-% it has revoked.
-
-% The \emph{public future} region relations satisfies the following properties:
-% \begin{mathpar}
-%   \inferrule{  (s,s') \in \phi_\pub \\
-%     (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H')}
-%   {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
-%   \and
-%   \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
-%   { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
-%   \and
-%   \inferrule{ }
-%   { \revoked \futurewk \revoked }
-% \end{mathpar}
-
-% Notice that the public future region relation is a subset of the
-% private future region relation, so anything you are allowed to do in
-% the public future region relation, the private future region relation
-% can do the same.
-
-% % The future world relation are really the ones from thm, the above
-% % are properties of them.
-% {\bf Future world relations}
-% There are two future world relations a \emph{public future world}
-% relation and a \emph{private future world} relation. In both future
-% world relations, we have extension ordering, so a future world $W'$ of
-% $W$ has at least the same regions as $W$, but $W'$ may have new regions
-% imposing new protocols on parts of the memory. The two future world
-% relations differ in the way existing regions are allowed to
-% evolve. The public future world relation uses the public future region
-% relation and the private future world relation uses the private future
-% region relation:
-% \begin{mathpar}
-%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-%     \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
-%   { W' \futurewk W }
-%   \and
-%   \inferrule{ \dom(W') \supseteq \dom(W)\\ 
-%     \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
-%   { W' \futurestr W }
-% \end{mathpar}
-
-% % Relate to STSs for high-level languages. Have the public-private
-% % transitions, but they are also used for more than this
-% \todo[inline]{Relate to STSs for high-level languages?}
-
-\subsubsection{World Satisfaction}
-A memory satisfies a world, written $\memSat{\ms}{W}$, if it can be partitioned
-into disjoint parts such that each part is accepted by an active (permanent or
-temporary) region.  Revoked regions are not taken into account as their memory protocols are no longer in effect.
-\begin{align*}
-  &\memSat{\ms}{W}
-    \text{ iff }\quad \left\{\begin{aligned}
-        &\exists P : \activeReg{W} \rightarrow \HeapSegments \ldotp \hs = \biguplus_{r\in\activeReg{W}}P(r) \text{ and } \\[-.35cm]
-        &\quad \forall r \in \activeReg{W} \ldotp\\
-        &\quad \quad \exists H,s \ldotp W(r) = (\_,s,\_,\_,H) \text{ and } \npair[n]{P(r)} \in H(s)(\xi^{-1}(W))\\
-      \end{aligned}\right.
-\end{align*}
-% LB: omitted the following for brevity:
-% Note that it is possible to state nonsensical worlds with conflicting
-% requirements for the same part of memory. There would, however, be no
-% memory that would satisfy such a world which for all intents and purposes would
-% make it discounted by the logical relation.
-
-\subsection{Logical Relation}
-\afterpage{
-\begin{figure}[htb]
-  \centering
-%  \begin{withmathindent}{0cm}
-    \begin{align*}
-      & \observations \; : \;  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
-      & \observations (W) \defeq 
-        \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
-        \begin{aligned}
-          & \forall \ms_f, \heap', i \leq n \ldotp  (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap') \Rightarrow \\
-          & \;\; %\Rightarrow
-            \exists W' \futurestr W, \ms_r, \ms' \ldotp \\
-          & \;\;\;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{ and } \heapSat[\hs']{n-i}{W'}
-        \end{aligned}\right\}\\
-      & \stdrr \; : \; \Worlds \monwknefun \UPred{\Regs} \\
-      & \stdrr(W) \defeq \left\{ \npair{\reg} \; \middle| \;
-          \forall r \in \RegName \setminus \{\pcreg\} \ldotp \npair{\reg(r)}
-          \in \stdvr(W) \right\} \\
-      & \\[-.2cm]
-      & \stder \; : \; \Worlds \nefun \UPred{\Words} \\
-      & \stder(W) \defeq \left\{ \npair{\pc} \; \middle| \;
-        \begin{multlined}
-          \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
-          \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W)
-        \end{multlined} \right\} \\
-      & \\[-0.2cm]
-      &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
-      &\stdvr(W)\defeq 
-        \begin{aligned}[t] & \{ \npair{i} \mid i \in \ints \} \union \\
-          &
-          \{ \npair{\stdcap[(\noperm,\gl)] } \}
-        \union \\
-        & \left\{
-            \npair{\stdcap[(\readonly,\gl)] } \mid \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-          \right\}
-        \union \\
-        & \left\{
-            \npair{\stdcap[(\readwrite,\gl)] } \; \middle| \;
-            \begin{array}{l}
-              \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-              \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
-            \end{array}
-          \right\} \union \\
-        & \left\{
-            \npair{\stdcap[(\readwritel,\gl)] } \;\middle| \;
-            \begin{array}{l}
-             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W)
-            \end{array}
-           \right\}
-        \union \\
-        & \left\{
-          \npair{\stdcap[(\exec,\gl)]} \; \middle| \;
-          \begin{array}{l}
-             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-          \end{array}
-           \right\}
-        \union \\
-        & \left\{
-            \npair{\stdcap[(\entry,\gl)]} \; \middle| \;
-            \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)
-          \right\}
-        \union \\ 
-        & \left\{
-          \npair{\stdcap[(\rwx,\gl)]}\; \middle| \; 
-          \begin{array}{l}
-            \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-            \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{ and }\\
-            \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-          \end{array}
-          \right\}
-        \union \\
-        & \left\{
-            \npair{\stdcap[(\rwlx,\gl)]} \; \middle| \;
-            \begin{array}{l}
-             \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and } \\
-             \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{ and }\\
-             \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W)
-            \end{array}
-          \right\}
-      \end{aligned}
-    \end{align*}
-%  \end{withmathindent}
-  \caption{The logical relation.}
-  \label{fig:logrel}            
-\end{figure}
-
-% \begin{figure}[htbp]
-%   \centering
-%   \begin{align*}
-%   &\iota_{\start,\addrend}^\pwl : \Regions\\
-%   &\iota_{\start,\addrend}^\pwl \defeq (\temp,1,=,=,H^\pwl_{\start,\addrend}) \\
-%   \\
-%   &H^\pwl : \Addrs^2 \fun \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-%   &H^\pwl_{\start,\addrend}\; s \; \hat{W} \defeq  \quad\left\{\npair{\hs} \; \middle| \;
-%     \begin{aligned}
-%       & n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }} \\
-%       &\forall \addr \in [\start,\addrend] \ldotp\\
-%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-%     \end{aligned}
-%         \right\}\\
-%   &\iota_{\start,\addrend}^{\nwl} : \Regions \\
-%   &\iota_{\start,\addrend}^{\nwl} \defeq (\temp,1,=,=,H^\nwl_{start,\addrend}) \\
-%   \\
-%   &H^\nwl : \Addrs^2 \fun \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-%   &H^\nwl_{\start,\addrend} \; s \;\hat{W} \defeq \quad\left\{\npair{\hs} \; \middle| \;
-%     \begin{aligned}
-%       &n = 0 \text{\footnotesize{ or }} (\dom(\hs) = [\start,\addrend] \text{\footnotesize{ and }}\\
-%       &\forall \addr \in [\start,\addrend] \ldotp \\
-%       & \;\; {\footnotesize\nonlocal{\ms(\addr)}} \text{\footnotesize{ and }}\\
-%       & \;\; \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W})))
-%      \end{aligned}
-%      \right\}
-% \end{align*}
-% \caption{Standard regions}
-% \label{fig:standard-regions}
-% \end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \begin{align*}
-   \readCond{}(\gl)(W) &=  \left\{ \npair{(\start,\addrend)} \; \middle| \;
-    \begin{array}{l}
-       \exists r \in \var{localityReg}(g,W) \ldotp \\
-\quad   \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
-    \end{array}
-    \right\}\\
-   \writeCond{}(\iota,\gl)(W) &=  \left\{
-    \npair{(\start,\addrend)}
-    \; \middle| \;
-    \begin{aligned}
-      & \exists r \in \var{localityReg}(g,W) \ldotp \\
-      & W(r) \text{ is address-stratified and} \\[-0.2cm]
-      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
-    \end{aligned} \right\}\\
-   \execCond{}(\gl)(W) &= 
-    \left\{
-      \npairP{(\var{P}, \start,\addrend)}
-     \; \middle| \;
-    \begin{aligned}
-      & \forall n' < n, W' \future W, a \in [\start',\addrend'] \subseteq [\start,\addrend], \perm \in \var{P} \ldotp \\
-      & \;\;\;\; \npair[n']{((\perm,\gl),\start',\addrend',\addr)} \in \stder(W')\\
-    \end{aligned} \right\} \\
-   \entryCond{}(\gl)(W) &= 
-    \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
-    \begin{array}{l}
-      \forall n' < n \ldotp \forall W' \future W \ldotp\\
-      \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
-    \end{array}
-    \right\} \\
-   \text{where } \gl = \local \Rightarrow \future = \futurewk \text{ and } \gl = \glob \Rightarrow \future = \futurestr \span
-   \end{align*}
-\caption{Permission-based conditions}
-\label{fig:perm-conds}
-\end{figure}
-\begin{figure}[htb]
-  \begin{align*}
-  \iota^\pwl_A, \iota^\nwl_A :{} & \Regions \\
-  \iota^\pwl_A \defeq{} & (\temp,1,=,=,H^\pwl_A) \\
-  \iota^\nwl_A \defeq{} & (\temp,1,=,=,H^\nwl_A) \\
-  \iota^{\nwl,p}_A \defeq{} & (\perma,1,=,=,H^\nwl_A)\\
-    H^\pwl_A :{} & \States \fun (\Wor \monwknefun \UPred{\HeapSegments})\\
-  H^\pwl_A (s)(\hat{W}) \defeq{} & \left\{\npair{\hs} \middle|
-    \begin{aligned}
-      &\dom(\hs) = A \text{ and } \\
-      &\forall \addr \in A \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
-    \end{aligned}
-        \right\} \union \{\npair[0]{\ms} \} \\
-    H^\nwl_A :{} & \States \fun (\Wor \monstrnefun \UPred{\HeapSegments})\\
-  H^\nwl_A (s)(\hat{W}) \defeq{} & \left\{\npair{\hs} \middle|
-    \begin{aligned}
-      &\dom(\hs) = A \text{ and } \\
-      &\forall \addr \in A \ldotp \\
-      &\quad \nonlocal{\ms(\addr)} \text{ and }\\ 
-      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
-    \end{aligned}
-        \right\} \union \{\npair[0]{\ms} \}
-\end{align*}
-
-  \caption{The standard permit write-local and no write-local regions.}
-  \label{fig:std-reg}
-\end{figure}
-}
-
-%subsection introduction
-% \dominique{I would try to motivate the LR defs by saying that the LR is
-%   constructed to express the FTLR and the FTLR says that if a piece of code only
-%   has access to values respecting the invariants in a world, then it will
-%   respect them itself too.} 
-The logical relation defines semantically when values, program counters, and configurations are capability safe.
-The logical relation is defined found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and we provide some explanations in the following paragraphs.
-For space reasons, we omit some definitions and explain them only verbally, but precise definitions can be found in the appendix.
-The logical relation is recursively defined, so we encourage first time readers to just read the section in its entirety and do a second read afterwards.
-
-First, the \emph{observation relation} $\observations$ defines what
-configurations we consider safe. A configuration is safe
-with respect to a world, when the execution of said configuration does not break
-the memory protocols of the world. Roughly speaking, this means that when the
-execution of a configuration halts, then there is a private future world that
-the resulting memory satisfies. Notice that failing is considered safe behavior.
-In fact, the machine often resorts to failing when an unauthorized access is
-attempted such as loading from a capability without read permission. This is
-similar to \citet{Devriese:2016ObjCap}'s logical relation for an untyped
-language but unlike typical logical relations for typed languages that
-require programs to not fail.
-
-The \emph{register-file relation} $\stdrr$ defines safe register-files as those that contain safe words (i.e.\ words in the $\stdvr$-relation defined below) in all registers but the $\pcreg$-register.
-The \emph{expression relation} $\stder$ defines what words are safe to use as a program counter.
-A word is safe to use as a program counter when it can be used to form a safe configuration by plugging it into the $\pcreg$-register of a safe register file (i.e.\ a register file in $\stdrr$) and pairing it with a memory satisfying the world.
-Note that integers and non-executable capabilities (e.g.\ $\readonly$ and $\entry$ capabilities) are considered safe program counters because when they are plugged into a register file and paired with a memory, the execution will immediately fail, which is safe.
-
-The \emph{value relation} $\stdvr$ defines when words are safe.
-\itoplassug{Explain why we use $\UPred{}$ with reference to the introduction of this section.}
-We make the value relation as liberal as possible by defining it based on the principle ``what is the most we can allow an adversary to use a capability for without breaking the memory protocols.''
-
-Non-capability data is always safe because it provides no authority.
-Capabilities give the authority to manipulate memory and potentially break memory protocols, so they need to satisfy certain conditions to be safe.
-In \figurename~\ref{fig:perm-conds}, we define such a condition for each kind of permission a capability can have.
-
-Capabilities with read permission cannot directly break memory protocols because they cannot make changes to the memory.
-However, a read capability could be used to read a capability that can break memory protocols.
-For this reason, a capability with read permissions is only safe when it can be used only to read safe capabilities.
-This is captured by the $\readCond{}$-condition. 
-More precisely, the condition requires the memory addressed by the capability to actually be governed by some region $W(r)$.
-This region should at least require all the memories accepted by the state interpretation function to contain safe words but stricter requirements may be imposed on the memories.
-This is expressed in terms of an upper bound on the possibly permitted memories defined in terms of a so called ``standard region''.
-In this case, it is a \tpwl{} $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
-The \tpwl{} only accept memories with address range $[\start,\addrend]$.
-More importantly, the \tpwl{}'s state interpretation function only accepts memory segments that only contain safe words.
-The relation between the actual region in the world and the upper bound is $W(r) \nsubsim[n] \iota^\pwl_{[\start,\addrend]}$ (defined in Appendix~\ref{app:logical-relation}).
-Essentially, the relation requires the two regions state transition systems to be the same.
-Further in the current state of region $W(r)$ and in any world, the state interpretation function should only allow memories that are also allowed in the standard region.
-
-The locality of capabilities play a role in whether or not they are safe.
-Generally speaking, global capabilities should only depend on permanent regions.
-This is expressed in $\readCond{}$ with $\var{localityReg}(\gl,W)$ which projects out all the regions the capability can depend on based on its locality.
-Specifically, if the capability is local, then all active regions (non-revoked) are project.
-On the other hand, if the capability is global, then only the permanent regions are projected.
-
-Capabilities with write permission can be used to write to memory, so to define when a capability with write permission is safe we ask ourselves what an adversary should be allowed to do with it.
-An adversary should at least be able to write safe words to memory.
-Safe words cannot be used to break memory invariants, so we will permit any safe word to be written.
-A capability with write permission\footnote{On the capability machine we consider, write permission always comes with read permission, so we could have made a stricter condition for write capabilities.}
-only allows you to write to memory not read from it, so we can allow anything to be written to memory as long as it cannot be read back and used to break memory protocols.
-For this reason, the condition on capabilities with write-permission is defined as a lower bound on what can be written where the lower bound is ``any safe word''.
-
-The condition on write capabilities is complicated by the fact that we have two flavours of write permission: write and write-local.
-A write-local capability can be used to write both local and global capabilities, so it is (at least) allowed to write any safe capabilities.
-Write capabilities, on the other hand, cannot write local capabilities, so we set the lower-bound as non-local safe words.
-
-The requirements on write-capabilities are captured by the $\writeCond{}$-condition, and it is defined in a similar fashion to the $\readCond{}$-condition.
-The regions the write-capability may depend on are projected from the world with the $\var{localityReg}$-function, and $\nsupsim[n-1]$ is used to relate the actual region to the lower bound.
-We need a different lower-bound depending on whether the permission of the capability is write or write-local, so $\writeCond{}$ is parameterized with the region it actually uses.
-If the capability has write-local permission, then we use the \tpwl{}.
-If capability only has write permission, then we use a \tnwl{} $\iota^\nwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}).
-In addition to the requirements of a \tpwl, the \tnwl{} requires all words in the accepted memory segments to be non-local.
-
-Finally, there is a technical requirement that the region must be \emph{address-stratified}.
-Intuitively, this means that if a region accepts two memory segments, then it must also accept every memory segment ``in between'' in the sense that it should be possible to come from one memory segment to the other.
-Our capability machine can only update one memory address at a time, so it should be accepted to construct a memory that is on its way to become the one we want to accept.
-Specifically, if two memory segments are accepted by a region, then the region must also accept every memory segment where each address contains a value from one of the two accepted memory segments.
-
-Due to the permission combinations on the capability machine we consider, a $\writeCond{}$ always comes with a $\readCond{}$ which creates a somewhat tight bound for the possible regions that make a capability safe.
-Further, we could have made the $\writeCond{}$ based on the fact that a writer permission always comes with a read permission, but we opted to make the logical relation as general as possible, so it can be reused in a setting with a richer permission hierarchy.
-
-The conditions $\entryCond{}$ and $\execCond{}$ are very similar. Both require
-that the capability can be safely jumped to. However, executable capabilities
-can be updated to point anywhere in their range, so they must be safe as a
-program counter (in the $\stder$-relation) no matter the current address.
-\itoplas{The range of an executable capability can also be reduced, so they must also be safe as program counter no matter what their range of authority is reduced to.}
-In contrast, enter capabilities are opaque and can only be used to jump to the
-address they point to.
-%
-\begin{toplas}%
-  This is why $\entryCond{}$ depends on the current address of the capability unlike for other types of capabilities.
-\end{toplas}
-They also change permission when jumped to, so we require
-them to be safe as a program counter after the permission is changed to $\exec$.
-% This is the only condition that depend on the current address, a, because this is the 
-% only kind of capability in which the current address cannot be changed.
-Because the capabilities are not necessarily invoked immediately, this must be
-true in any future world, but it depends on the capability's locality which
-future worlds we consider. If it is global, then we require safety as a program
-counter in \emph{private} future worlds (where temporary regions may be
-revoked). For local capabilities, it suffices to be safe in \emph{public} future
-worlds where temporary regions are still present.
-
-In the technical appendix, we prove that safety of all values is
-preserved in public future worlds, and that safety of global values is
-also preserved in private future worlds:
-\begin{lemma}[Double monotonicity of value relation]~
-  \begin{itemize}
-  \item If $W' \futurewk W$ and $\npair{w} \in \stdvr(W)$, then $\npair{w} \in
-    \stdvr(W')$.
-  \item If $W' \futurestr W$ and $\npair{w} \in \stdvr(W)$ and $w =
-    \stdcap[(\perm,\glob)]$ (i.e.\ $w$ is a global capability), then $\npair{w}
-    \in \stdvr(W')$.
-  \end{itemize}
-\end{lemma}
-
-\begin{toplas}
-  In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
-  This indicates that our logical relation should imply the same, namely that capabilities with write-local permission are local.
-  \begin{lemma}[Stack capabilities are local]~
-    \label{lem:wl-local}
-    \begin{itemize}
-    \item If $\memSat{\ms}{W}$ and $\npair{\stdcap} \in \stdvr(W)$ and $\start
-      \leq \addrend$ and $\perm \in \{\rwlx,\rwl\}$, then $\gl = \local$.
-    \end{itemize}
-  \end{lemma}
-  In our definition of worlds, nothing prevents a world from having regions that are overlapping.
-  This might seem like an issue the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
-  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness condition on the world as it prevents the regions from overlapping.
-  The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
-\end{toplas}
-
-\subsection{Capability Machine Safety}
-With the logical relation defined, we can now state the fundamental theorem of
-our logical relation: a strong theorem that formalizes the guarantees offered by
-the capability machine. Essentially, it says a capability that only grants safe
-authority is capability safe as a program counter.
-\begin{theorem}[Fundamental Theorem]
-  \label{thm:ftlr}
-  If one of the following holds:
-  \begin{align*}
-      & \bullet
-        \begin{aligned}[t]
-        &\perm = \exec \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
-      \end{aligned} \\
-    & \bullet 
-      \begin{aligned}[t]
-        &\perm = \rwx \text{ and } \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
-        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
-      \end{aligned} \\
-    & \bullet 
-      \begin{aligned}[t]
-        &\perm = \rwlx \text{ and }\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{ and }\\
-        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
-      \end{aligned}
-  \end{align*}
-  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
-\end{theorem}
-\begin{toplas}
-  \begin{proof}[Proof sketch]
-    Induction over $n$. By definition of $\stder(W)$, show
-    \[
-      \npair[n']{(\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms)}
-      \in \observations(W)
-    \]
-    assuming $n' \leq n$, $\npair[n']{\reg} \in \stdrr(W)$, and
-    $\heapSat[\ms]{n'}{W}$. By definition of $\observations$ let $\ms_f$,
-    $\mem'$, and $i \leq n'$ be given and for
-    $\Phi = (\reg\update{\pcreg}{((\perm,\gl),\start,\addrend,\addr)},\ms \uplus
-    \ms_f)$ assume $\Phi \step[i]
-    (\halted,\mem')$ and show there exists $W' \futurestr W$ that part of
-    $\mem'$ satisfies. First observe that $i \neq 0$ as $\Phi$ is a non-halted
-    configuration, so $\Phi$ takes at least one step, i.e.\ $\Phi \step
-    \Phi' \step[i-1] (\halted,\mem')$.
-
-    The rest of the proof considers the different possibilities for the step $\Phi \step
-    \Phi'$, i.e. it considers each of the instructions that could have been executed. For
-    each of these cases, we argue that (1) $\Phi'$ is consistent with the world
-    in the sense that the register-file and memory still respect the world and
-    that (2) the rest of the execution respects the world. Depending on where
-    the $\pcreg$ in $\Phi'$ comes from, the second result is proven in one of
-    two ways. If the step to $\Phi'$ was a jump, then the new $\pcreg$ is one of
-    the safe values in $\Phi$'s registers, and the value relation is used to
-    argue that the jump is safe. On the other hand, if the $\pcreg$
-    was just incremented, then we can apply the induction hypothesis.
-
-    In order to argue (1), we look at the possible states for $\Phi$ according to the operational semantics.
-    If we consider the memory in $\Phi'$, then it either (a) remains unchanged or (b) one address has been updated and the register-file contains an appropriate capability for writing.
-    The latter occurs when the executed instruction is $\texttt{store}$.
-    Otherwise we are in the former case.
-    In case (a), the memory hasn't changed, so we conclude that the memory still respects the world simply by downwards closure of memory satisfaction.
-    In case (b), we use an auxiliary lemma that uses the safety of the write capability used by the $\texttt{store}$ instruction to show that the updated memory satisfies the world.
-    To show that the updated register-file is safe, we consider the changes made to it by all instructions in separate lemmas and show that they all preserve safety of the register file.
-
-    The complete proof can be found in the technical appendix~\citep{technical_appendix}.
-  \end{proof}
-\end{toplas}
-The permission-based conditions of Theorem~\ref{thm:ftlr} make sure that the
-capability only provides safe authority in which case the capability must be in
-the $\stder$ relation, i.e. it can safely be used as a program counter in an
-otherwise safe register-file.
-
-The Fundamental Theorem can be understood as a general expression of the guarantees offered by the capability machine which is an instance of a general property called capability safety~\citep{Devriese:2016ObjCap,Maffeis2010OC}.
-The theorem says that an arbitrary capability $((\perm,\gl),\start,\addrend,\addr)$ is safe as a program counter without making any assumptions about what program it points to (the only assumptions we have are about the read or write authority that it carries).
-As such, the theorem expresses the capability safety of the machine which guarantees that \emph{any} instruction is safe to execute and will not be able to go beyond the authority of the values it has access to.
-We demonstrate this in \sectionname~\ref{sec:examples} where Theorem~\ref{thm:ftlr} is used to reason about capabilities that point to arbitrary instructions.
-The relation between Theorem~\ref{thm:ftlr} and local-state encapsulation and control-flow correctness will also be shown by example in \sectionname~\ref{sec:examples} as the examples depend on these properties for correctness.
-
-% See the technical appendix~\citep{technical_appendix} for a detailed
-% proof (by induction over the step-index $n$) of the theorem.
-
-% LB: following omitted for brevity:
-% From another point of view, the Fundamental Theorem also serves as a sanity
-% check of the logical relation as it guarantees a large class of
-% capabilities actually inhabits it.
-
-% The FTLR does not mention integers and non-executable capabilities because they are trivially safe (it is safe to fail)
-
-\itoplassug{Suggestion: Add a small section about reasoning about programs.
-Add the ``scall works'' and ``malloc works'' lemmas and describe how they are used.
-This section could also point out the things one have to argue about that are hidden in similar high-level proofs (e.g., the return pointer valid when we return corresponding to restoring memory invariants of caller.).}
-
-\begin{toplas}
-  \section{Malloc}
-  \label{sec:malloc}
-  In the examples presented in \sectionname~\ref{sec:examples}, we will assume the existence of a trusted malloc routine, so that both the trusted code and the adversary are able to allocate new memory.
-  Malloc is considered part of the trusted computing base as mentioned in \sectionname~\ref{sec:capab-mach-with}.
-  This is unavoidable: if we cannot trust malloc, then we cannot use the memory it allocates as we have no idea whether it is aliased by some untrusted program.
-
-  Our semantic model is not specific to a particular implementation of malloc, so rather than providing the implementation we provide a malloc specification.
-  The specification expresses standard expectations about the behaviour of malloc, but making the specification realistic requires some of the technical machinery from the logical relation.
-  As such, this section is a bit technical and can safely be skipped on first read.
-  Definition~\ref{spec:malloc} is the malloc specification.
-  Following the definition, we provide an informal description of the definition with references to each of the items in the definition.
-
-\begin{definition}[Malloc Specification]
-  \label{spec:malloc}
-  $c_\malloc$ satisfies the specification for malloc iff the following
-  conditions hold:
-  \begin{enumerate}
-  \item \label{item:malloc-spec:eg} $c_\malloc = ((\entry,\glob),\_,\_,\_)$
-  \item \label{item:malloc-spec:reg} There exists a $\iota_{\malloc,0}$ such that
-    \begin{enumerate}
-    \item \label{item:malloc-spec:perm} $\iota_{\malloc,0}.v = \perma$
-    \item \label{item:malloc-spec:independent} For all $\iota' \futurestr \iota_{\malloc,0}, W,i$ with
-      $W(i)=\iota'$, we have that
-      \begin{equation*}
-        \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
-      \end{equation*}
-    \item \label{item:malloc-spec:correct-exec} For all $\Phi \in \ExecConfs$, $\ms_{\var{footprint}}, \hsframe \in
-      \HeapSegments$, $i, n, \size \in \nats$, $w_{\var{ret}} \in \Words$,
-      $\iota_\malloc \futurestr \iota_{\malloc,0}$, we have that
-      \[
-        \begin{array}{l}
-          \text{If }
-          \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land{} \\
-          \quad\memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land{} \\
-          \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
-          \text{Then, } \\
-          \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
-          \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
-          \qquad \quad \Phi \step[j] \Phi' \land{} \\
-          \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land{}\\
-          \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land{} \\
-          \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land{} \\
-          \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land{} \\
-          \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land{} \\
-          \qquad \quad \size - 1 = e'-b'
-        \end{array}
-      \]
-    \item \label{item:malloc-spec:fail} For all $\Phi \in \ExecConfs$,
-      \[
-        \begin{array}{l}
-          \text{If } (\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
-          \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
-          \text{Then } \exists j \in \nats \ldotp \Phi \step[j] \failed
-        \end{array}
-      \]
-    \end{enumerate}
-  \end{enumerate}
-\end{definition}
-
-  We require a global capability for invoking malloc (because if the capability were local, then a program with access to malloc would have to give up this access when invoking untrusted code, \ref{item:malloc-spec:eg}).
-  The capability is assumed to have enter permission (\ref{item:malloc-spec:eg}), so malloc can protect its internal state even when the capability is shared with untrusted code.
-
-  In addition to these syntactic requirements, we also specify standard behavioural expectations of malloc.
-  Intuitively, we require that when malloc is invoked with a length argument, then it will return a capability for a fresh piece of memory of that size.
-  It should be fresh in the sense that malloc has not already allocated any of that memory and will not do so in the future.
-  Also, when invoked with a nonsensical length argument such as a negative integer or a capability, malloc should simply fail.
-  However, formulating these requirements is harder than one might expect.
-  The problem is that a realistic implementation of malloc needs to rely on internal state that changes after every invocation and relies on invariants on that state.
-  We can only expect that malloc behaves according to its specification if its internal state satisfies the current state of the invariants in an executing system.
-  We express this in terms of the semantic model from \sectionname~\ref{sec:logical-relation}.
-
-To allow malloc implementations to rely on internal state and invariants for
-that state, we assume an initial region for malloc (\ref{item:malloc-spec:reg}). The region is assumed to be
-permanent (since safety of the global malloc capability will depend on its
-presence, \ref{item:malloc-spec:perm}).
-% DOMI: I don't understand what the following uncommented text means
-% This also means that the malloc specification should guarantee that
-% use of malloc is safe at any point of execution no matter who has invoked it. To
-% this end, the requirement on the initial malloc region should also hold true in
-% any private future world.
-% priv future means that malloc should not depend on local capabilities?
-Furthermore, we want to express that malloc does not depend on any other memory
-than its own internal state. This is expressed by a restriction on the malloc
-region's state interpretation function which (as explained in
-\sectionname~\ref{subsec:worlds}) defines what memory segments it permits in a
-given world. We require that the accepted memory segments only depend on the
-presence of the malloc region itself, i.e.\ in any world the same memory
-segments are accepted if we remove all regions except the malloc region. This
-property should continue to hold throughout execution, so it must hold true for
-any private future region of the initial malloc region (\ref{item:malloc-spec:independent}).
-
-The malloc specification also dictates what malloc should do when invoked in a
-memory with its internal state valid according to the malloc region (some future
-evolution of the initial malloc region). If malloc is invoked with an invalid
-length argument (that is, a negative integer or a capability), then we simply
-require malloc to fail (\ref{item:malloc-spec:fail}). This part of the
-specification does not actually rely on the malloc region: for simplicity, we
-assume malloc does not need its internal state to check the argument. When
-malloc is invoked with a valid length (\ref{item:malloc-spec:correct-exec}),
-then it should return a fresh memory segment of the correct length. This segment
-is required to come from the footprint of malloc, i.e.\ the memory owned by the
-malloc region before the call. After malloc returns, we require the malloc
-region to have evolved (according to the public future world relation) to a new
-state where the new footprint is disjoint from the allocated memory. This
-implies that future invocations of malloc can never return previously-allocated
-memory.
-
-For convenience, we also require that malloc returns the non-argument registers
-of the register-file unchanged after the call. This allows the caller to keep
-private capabilities in the register file, without having to protect them by
-storing them in a private stack frame.
-
-As described previously, the specification of malloc ensures that malloc has no
-capabilities pointing \textit{out} of malloc. It does not, however, say anything
-about capabilities that point \textit{in} to malloc. If we want to be able to
-trust malloc, we obviously cannot allow arbitrary capabilities to point in to
-it. We have chosen to keep the malloc specification simple and let this
-assumption be in the lemmas that use malloc. It is sufficient for these lemmas
-to require that there are no outside capabilities for malloc in the initial
-configuration as capabilities cannot appear out of thin air, and the malloc
-specification makes sure that capabilities are not leaked.
-
-Malloc should not just be available to trusted programs but also to possibly
-malicious programs. This is safe as it follows from the specification that the malloc
-capability is always safe in a world with the malloc region:
-\begin{lemma}[Malloc is safe to pass to adversary]
-  \label{lem:malloc-in-vr}
-  For all $c_\malloc$ that satisfies the specification for malloc with region $\iota_{\malloc,0}$,  if $W(r) \futurestr \iota_{\malloc,0}$, then
-  $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
-\end{lemma}
-
-The reason we allow trusted code and adversary to invoke malloc is just to make
-our work more realistic, but we are otherwise not interested in its details. As
-such, we do not give a malloc implementation. We are, however, confident that it
-is possible to make an implementation of malloc that satisfies the malloc
-specification in Definition~\ref{spec:malloc}. There are in fact two
-simplifications in our system that makes things easier: First, we do not consider
-deallocation of memory which means that the data structure malloc uses to keep
-track of free memory does not have to handle reclaimed memory. Second, the
-malloc specification does not permit malloc to run out of memory and thus refuse
-allocation. This is possible on our simple capability machine because it has an
-infinite address space. An initial capability with an infinite range of
-authority would of course need to be part of malloc, but it could also double as
-the data structure that keeps track of free memory. % \itoplassug{If we have
-  % references for malloc being difficult, we could add that to the text.}
-%\itoplassug{Add the assumption that the malloc capability in examples
-%  satisfies specification.}
-
-\section{Reusable macro instructions}
-\label{sec:macros}
-%\itoplassug{Add examples of macros to this section to make it clear what we mean by macros.\\}
-%\itoplassug{Reviewer C, ESOP, asks for high-level specifications for the macros in the style of malloc.
-I am not sure we want to actually include this in the submission, so it is maybe not here we should put our attention.
-(see review in tex comment)\\}
-% Finally, as already mentioned above, the approach relies on `call` and `scall` macro-instructions that are only outlined in the submission, and described by very complicated pseudocode and English prose in the companion technical appendix.
-I still miss a fully formal description of 1- what instructions the macros expand to, exactly, and 2- what contract they satisfy (in the style of the contract for malloc in section 2 of the technical appendix).
-With the calling convention and logical relation defined, we would like to show its usefulness by proving
-the correctness of a series of examples that rely
-on well-bracketedness and local-state encapsulation and use the calling
-convention to enforce these properties. However, the programs that
-run on our capability machine are assembly programs. Program examples that would
-be small in a high-level language become big and unintelligible at this
-low level. Thus to make our program examples intelligible, we introduce a
-series of low-level abstractions in the form of macros.
-We define a number of reusable macros capturing the calling convention as well
-as other conveniences. The macros that utilize the stack assume that it is
-available in register $r_\stk$.
-\itoplassug{Lau, suggestion: should we mention temporary registers? They are
-  used in the macros, but it seems like a minor technical detail.}
-
-The macro \texttt{scall
-  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures the
-parts of the calling convention related to actually transferring control to
-adversarial code. Specifically, it pushes the contents of the private registers,
-$\overline{r_{\var{priv}}}$, to the stack and pushes the ``restoration'' code to
-the stack. The restoration code is executed as the first thing upon return; it
-restores the stack pointer and the old program counter. After the restoration
-code is pushed to the stack, \texttt{scall
-  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} adjusts the pc
-to point to the first instruction after the jump and pushes it to the stack. The
-\texttt{scall}macro constructs a protected return pointer from a stack pointer by
-adjusting it to point to the first instruction of the return code and
-encapsulating it by restricting its permission to $\entry$. Next, the \texttt{scall}
-macro reduces the range of authority of the stack pointer to the unused part of
-the stack and clears it(as discussed in
-Section~\ref{sec:stack-and-return-pointer}). Finally, the \texttt{scall} macro clears
-non-argument registers and jumps to $r$. Upon return after the restoration code
-on the stack has been executed, the \texttt{scall} macro pops the restoration code from
-the stack and restores the private state by popping it from the stack to the
-appropriate registers. \figurename~\ref{fig:scall-impl} displays the
-implementation of \texttt{scall} and the restoration code used by
-\texttt{scall}. The implementation of \texttt{scall} uses some of the macros we
-present next.
-
-\begin{figure}[htb]
-  \begin{toplas}
-    \centering
-    \begin{minipage}[t]{.6\linewidth}
-\begin{lstlisting}
-// Push the private registers to the stack.
-  push $r_{\var{priv},1}$
-  ...
-  push $r_{\var{priv},n}$
-// Push the restoration code to the stack.
-  push $\encode(i_1)$
-  push $\encode(i_2)$
-  push $\encode(i_3)$
-  push $\encode(i_4)$
-// Push the old pc to the stack.
-  move r_t1 pc
-  lea r_t1 $\var{off}$ 
-  push r_t1
-// Push the stack pointer to the stack.
-  push r_stk
-// Set up the protected return pointer.
-  move r_0 r_stk
-  lea r_0 $\var{off_\var{rec}}$
-  restrict r_0 $\encodePermPair((\local,\entry))$
-// Restrict the stack capability to the unused part of the stack.
-  geta r_t1 r_stk
-  plus r_t1 r_t1 1
-  getb r_t2 r_stk
-  subseg r_stk r_t1 r_t2
-// Clear the unused part of the stack.
-  mclear r_stk
-// Clear non-argument registers.
-  rclear $R$
-  jmp r
-return:
-// Pop the restore code.
-  pop r_t1
-  pop r_t1
-  pop r_t1
-  pop r_t1
-// Pop the private state into approriate registers.
-  pop $r_{\var{priv},1}$
-  ...
-  pop $r_{\var{priv},n}$
-\end{lstlisting}      
-    \end{minipage}
-    \begin{minipage}[t]{.29\linewidth}
-\begin{lstlisting}[title=\footnotesize{The restoration code used in \texttt{scall}.}, captionpos=b,frame=single,numbers=none]
- $i_1$ = move r_t1 pc
- $i_2$ = lea r_t1 $\var{off}_\var{stk}$
- $i_3$ = load r_stk r_t1
- $i_4$ = pop pc
-\end{lstlisting}
-    \end{minipage}%
-    \caption{Implementation of \texttt{scall} $r(r_{\var{args},1},\dots,
-      r_{\var{args},m},r_{\var{priv},1},\dots, r_{\var{priv},n})$. The
-      restoration code is presented in the top right corner. The
-      variable $\var{off}_\var{ret}$ is the offset to the label \texttt{return}, and
-      $\var{off_\var{rec}} = -5$ which is the offset to the first instruction of
-      the activation record. The set $R = \RegName \setminus
-      \{\texttt{pc},\texttt{r\_stk},\texttt{r\_0},r,r_{\var{args},1},\dots,r_{\var{args},m}\}$.
-    The variable $\var{off}_\var{stk} =5$ which is the offset to the address
-    where the old stack pointer is stored on the stack.}
-    \label{fig:scall-impl}
-  \end{toplas}
-\end{figure}
-
-The macro \texttt{mclear $r$} clears all the memory addresses that the
-capability in register $r$ has authority over. It is used by \texttt{scall} to
-clear the unused part of the stack before control is transferred. It should also
-be used to clear the stack before returning to adversarial code. Similarly, the
-macro \texttt{rclear $R$} clears all the registers in the set
-$R$. It is also used by \texttt{scall}, and it should also be used before
-returning to adversarial code.
-
-The macros \texttt{prepstack $r$} and \texttt{reqglob
-    $r$} are the last macros related to the calling convention. The former,
-\texttt{prepstack} should be used when one receives a stack from an unknown
-source. The \texttt{prepstack} macro first ensures the stack capability has permission
-read/write-local/execute. Then, it adjusts the stack capability, so it
-follows the convention for the stack\footnote{The stack capability should always
-point to the topmost word on the stack. A stack received from an unknown source
-can be treated as empty, so the stack capability should point just outside its
-range of authority.}. The other macro, \texttt{reqglob}, ensures that the
-capability in register $r$ is global. This macro should be used on callbacks
-received from unknown sources to ensure that they cannot be derived from the
-stack pointer.
-
-The remaining macros are not directly related to the calling convention, but
-they help making the examples in \sectionname~\ref{sec:examples} intelligible.
-The macros \texttt{push $r$} and \texttt{pop $r$} respectively add and remove
-elements from the stack. The macro \texttt{fetch $r$ $\var{name}$} fetches the
-capability related to $\var{name}$ from the linking table and stores it in
-register $r$. The macro \texttt{malloc $r$ $n$} invokes malloc with size
-argument $n$. The \texttt{malloc} macro assumes that a capability for malloc
-resides in the linking table and is basically a fetch of the malloc capability
-followed by a setup of a return pointer. Finally, the macro \texttt{crtcls
-  $\overline{(x_i,r_i)}$ $r$} allocates a closure where $r$ points to the
-closure's code and a new environment is allocated (using malloc) and the
-contents of $\overline{r_i}$ are stored in the environment. In the code referred
-to by $r$, an implicit load from the environment happens when an instruction
-refers to $x_i$.
-
-The Appendix contains the implementation of all the macros used in
-\texttt{scall}. The technical appendix~\citep{technical_appendix} contains more
-detailed descriptions of all the macros as well as all implementations.
-%\itoplassug{Suggestion: Add the call implementation (and possibly other macros?) in the above?}
-We stress that the macros correspond to series of instructions as seen in
-\figurename~\ref{fig:scall-impl}; the macros
-are introduced for intelligibility. The examples of
-\sectionname~\ref{sec:examples}, the program examples are stated using the
-macros, but the proofs work on the expanded macros.
-\end{toplas}
-
-\begin{toplas}
-\section{Reasoning about programs on a capability machine}
-\label{sec:reasoning}
-% \itoplassug{
-%   \begin{itemize}
-%   \item A section with the lemmas we use for reasoning about programs
-%     \begin{itemize}
-%     \item lemma 58, scall works. Include description of how it handles the
-%       common things in call and leaves the argumentation for callee code and
-%       continuation.
-%     \item lemma 59, malloc works. If this is included, we should probably also
-%       include the malloc specification.
-%     \item lemma 60, create closure works.
-%     \item Include a description about how the FTLR is used to reason about
-%       unknown code.
-%     \item Mention that this is a step on the way towards having a program logic,
-%       and then in the discussion discuss why a program logic for this language
-%       is not what we need.
-%     \end{itemize}
-%   \end{itemize}
-% }
-There are many details to get right when programming in assembly.
-These details carry over to proofs about assembly programs, so many of the proofs in \sectionname~\ref{sec:examples} about example programs are a bit cumbersome.
-It is especially annoying when the same line of reasoning is applied in multiple places.
-To mitigate this, we have define a number of lemmas that capture common reasoning patterns in these proofs.
-In this section, we present the most central lemmas used to reason about assembly programs.
-
-% Anti-reduction lemma
-Our capability machine only allows the final memory of an execution to be observed, so naturally the correctness lemmas we prove in \sectionname~\ref{sec:examples} are statements about the memory in the halted configuration.
-In order to prove a property about part of the final memory, we create a world with a region the ensures the desired property and show that the initial configuration is in the $\observations$-relation w.r.t.\ that world.
-The region that ensures the property must be permanent, so it is not revoked during execution.
-The $\observations$-relation says that if the initial configuration halts successfully, then the memory in the halting state must still respect a private future world of the initial world.
-It is, however, bothersome and error prone to try to argue about the entire execution in one go.
-Instead we want to reason modularly in the sense that we only want to reason about parts of the execution at a time.
-To this end, we prove an anti-reduction lemma that essentially says that if we can show that an initial configuration $\Phi$ steps to a configuration $\Phi'$ and $\Phi'$ is in the $\observations$-relation, then $\Phi$ must also be in the observation relation.
-\begin{lemma}[Anti-reduction for $\observations$]
-\label{lem:anti-red-obs}
-  \begin{align*}
-    & \forall n, n',i,\reg, \reg', \ms, \ms', \ms_r, W, W' \ldotp\\
-    & \quad n' \geq n - i \land W' \futurestr W \land \\
-    & \quad (\forall \ms_f \ldotp (\reg,\ms \uplus \ms_r \uplus \ms_f) \step[i] (\reg',\ms' \uplus \ms_r \uplus \ms_f)) \land \\
-    & \quad \npair[n']{(\reg',\ms')} \in \observations(W') \\
-    & \qquad \Rightarrow \npair{(\reg,\ms \uplus \ms_r)} \in \observations(W)
-  \end{align*}
-\end{lemma}
-
-% FTLR
-We use the anti-reduction lemma when we reason about the execution of known code.
-When we want to reason about unknown code, the anti-reduction lemma does not apply because we do not know what instructions are being executed.
-This is where the FTLR (Theorem~\ref{thm:ftlr}) comes into play.
-As a reminder, the FTLR says that if a capability only has access to safe values with respect to a world, then it is safe to use the capability for execution in the same world.
-The unknown code we consider will be assumed to only have access to safe values which typically means it has access to a linking table with a malloc capability and otherwise consists of instructions.
-These assumptions allow us to use the FTLR to reason about the unknown code as malloc is a safe value (cf.\ Lemma~\ref{lem:malloc-in-vr}) and instructions are integers; integers are always safe values.
-It is important to remember that safety is relatively to the semantic model of a memory and not the actual memory itself.
-For this reason, the assumptions we make on unknown code must be expressed as a region in a world.
-
-In order to argue that a specific configuration is safe, we need to argue that the configuration is safe with respect to the world.
-That is, we need to argue that the memory satisfies the world and the register-file is in the $\stdrr$-relation.
-For instance in the case where the untrusted code assumes control first, we will use the FTLR to show that the capability for the unknown code can be used for execution.
-The unknown code will get access to some known, trusted code through a capability in the initial register-file, so we will have to argue that the known code is safe.
-We argue about the known code with Lemma~\ref{lem:anti-red-obs}.
-
-% Many details to get right when reasoning about low-level programs.
-Another common pattern in proofs about programs on our capability machine concerns the use of \texttt{scall}.
-The following lemma captures the commonalities of reasoning about programs using \texttt{scall}.
-For instance, setting up the local stack frame and constructing stack pointers and protected return pointers for the callee always amount to the same line of reasoning which is captured by the lemma.
-From another point of view, it can be seen as a specification for \texttt{scall}.
-\begin{lemma}[$\mathtt{scall}$ works]
-  \label{lem:scall-works}
-  For all $n \in \nats$, $\ms, \ms_\stk, \ms_\var{unused}, \ms_f \in \MemSegments$, $W
-  \in \Worlds$, $\reg \in \Regs$, $r,\overline{r_{\mathit{arg}}},
-  \overline{r_{\mathit{priv}}} \in \RegName$, and $c_\var{next} \in \Caps$,
-  if
-  \begin{enumerate}
-  \item \label{item:scall:memsat} $\memSat[n]{\ms}{\revokeTemp{W}}$ 
-  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
-  \item \label{item:scall:stack} $(\reg,\ms) \text{ is looking at }
-    \mathtt{scall}\;r(\overline{r_{\mathit{arg}}},
-    \overline{r_{\mathit{priv}}}) \text{ followed by }
-    c_{\mathit{next}}$
-  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
-  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} For all  $\ms_\var{rec},
-    \ms_\var{unused}', \ms'' \in \MemSegments$, $W' \in \Worlds$, $c_\var{ret},
-    c_\stk \in \Caps$, and $\reg' \in \Regs$, if
-    \begin{itemize}
-    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{rec}}
-      \uplus \ms_{\mathit{unused}}')$,
-    \item $W' =
-      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
-    \item $\memSat[n-1]{\ms''}{W'}$
-    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
-    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
-      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
-      \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r \mapsto \reg(r)]$ 
-    \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
-    \item $\npair[n-1]{c_\stk} \in \stdvr(W')$
-    \end{itemize}
-
-    then we have $\npair[n-1]{(\reg',\ms'')} \in \observations(W')$
-  \item \label{item:scall:hyp-cont} \emph{Hyp-Cont} For all $n' \in \nats$, $W''
-    \in \Worlds$, $\ms'', \ms_\var{unused}''$, and $\reg' \in \Regs$, if
-    \begin{itemize}
-    \item $n' \leq n-2$
-    \item $W'' \futurewk \revokeTemp{W}$
-    \item $\memSat[n']{\ms''}{\revokeTemp{W''}}$ 
-    \item for all $r$, we have:
-      \begin{equation*}
-        \reg'(r)
-        \begin{cases}
-          = c_{\mathit{next}} &\text{ if } r = \pcreg\\
-          = \reg(r)&\text{ if } r \in \overline{r_{\mathit{priv}}}\\
-          \in \stdvr(\revokeTemp{W''}) &\text{ if $\reg'(r)$ is a global capability and } r \not\in \{\pcreg,\overline{r_{\mathit{priv}}}, r_\stk\}
-        \end{cases}
-      \end{equation*}
-    \item $\reg' \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}''$ unused}$
-    \end{itemize}
-
-    then we have $\npair[n']{(\reg',\ms'' \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}}'')} \in \observations(W'')$
-  \end{enumerate}
-  Then 
-  \begin{itemize}
-    \item $\npair{(\reg,\ms \uplus \ms_f \uplus \ms_\stk \uplus \ms_{\var{unused}})} \in \observations(W)$
-  \end{itemize}
-
-  % Note: $k_{\mathit{act-rec}}$ is the (standard) size of an activation record (2
-  % words for stack pointer and return pointer and some fixed number of words for
-  % the standard return instructions).
-\end{lemma}
-\begin{toplas}
-  The \texttt{scall} lemma is stated in terms of a number of auxiliary definitions found in the appendix.
-  The region $\iota^{sta}(\gl,\ms)$ is a static region with locality $\gl$.
-  It is static in the sense that it only accepts the memory segment $\ms$ (\appendixname~\ref{app:awkward-example}).
-  The function $\revokeTemp{} : \Worlds \fun \Worlds$ yields the input world but with all the temporary regions replaced with revoked regions (\appendixname~\ref{app:logical-relation}).
-  The definition of \textit{$(\reg,\ms)$ is looking at $[i_0\dots i_n]$ followed by $c_{\var{next}}$} is visualized in \figurename~\ref{fig:def-intuition-looking-at}; it requires the capability in $\pcreg$ of $\reg$ to point to the first address of the instructions $[i_0\dots i_n]$ in $\ms$ and $c_{\var{next}}$ to point to the address immediately after the instructions.
-  The definition of \textit{$\reg$ points to stack with $\ms_\stk$ used and $\ms_{\stk,\var{unused}}$ unused} is visualized in \figurename~\ref{fig:def-intuition-points-to-stack}; it requires $\ms_\stk$ and $\ms_{\stk,\var{unused}}$ to be adjacent and continuous memory segments with the local $\rwlx$-capability in \texttt{r\_stk} of $\reg$ governing the two memory segments and pointing to the top most address of $\ms_\stk$ (\appendixname~\ref{app:programs-definitions}).
-\end{toplas}
-
-Roughly, the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}) states that an invocation of scall is safe if \texttt{scall} is executed in a reasonable state (\ref{item:scall:memsat}-\ref{item:scall:stack}), the callee is safe to execute (\ref{item:scall:hyp-callee}), and returning to the code after the scall is safe (\ref{item:scall:hyp-cont}).
-
-In the common case, \texttt{scall} is used to reason about untrusted code that we only have general assumptions about.
-As explained previously, we use the FTLR (Theorem~\ref{thm:ftlr}) to argue about unknown code; this is no exception.
-That is, when we argue that we assumption~\ref{item:scall:hyp-callee} in the \texttt{scall} lemma satisfied, we use the FTLR along with the assumptions we have (e.g.\ linking table, malloc capability etc.).
-In Hyp-Callee, we assume the memory is safe, so it suffices to show that the register-file contains safe values, which amounts to showing that the arguments in the call are safe.
-
-By using Lemma~\ref{lem:scall-works} to reason about \texttt{scall}, the proofs become agnostic to the actual implementation of \texttt{scall}.
-In other words, should we change the implementation of \texttt{scall}, then we just need to prove that Lemma~\ref{lem:scall-works} holds for the new implementation in order to get that all our results still hold true.
-
-\begin{figure}[bt]
-  \centering
-  \begin{subfigure}{0.43\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-      \scope
-      \clip (-.1,-.1) rectangle (4.6,11.1);
-      \fill[fill=white] (0,0) rectangle (4.5,11);
-      \fill[color=gray!20] (0,0) rectangle (4.5,4);
-      \fill[color=gray!20] (0,9.5) rectangle (4.5,11);
-      \draw[color=gray!50] (0,5.5) rectangle (4.5,6);
-      \draw[color=gray!50] (0,7) rectangle (4.5,7.5);
-      \draw (0,0) -- (0,11);
-      \draw (4.5,0) -- (4.5,11);
-      \draw (0,4) -- (4.5,4);
-      \draw (0,9.5) -- (4.5,9.5);
-      \endscope
-      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-      \draw (-1.5,0) node {};
-      \draw (-1.5,15) node {};
-
-      \node at (8.5,7.5) { $\reg$ };
-      \begin{scope}
-        \clip (7.9,4.5) rectangle (9.1,7.1);
-        \foreach \x in {4,5,...,6}
-        {
-          \draw (8,\x) rectangle (9,\x + 1);
-        };
-      \end{scope}
-      \node[anchor=west] at (9.125,6.5) {\footnotesize\texttt{pc}};
-
-      \node at (2.25,7.25) { \footnotesize{$i_0$} };
-      \node at (2.25,6.68) { \footnotesize{$\vdots$} };
-      \node at (2.25,5.75) { \footnotesize{$i_n$} };
-      \node[rotate=0] at (2.25,8.5) {\huge{$\ms$}};
-
-      \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.59,mirror},yshift=0pt]
-      (4.6,4) -- (4.6,9.5) node (cstk) [black,midway,xshift=0.8cm] {};
-      \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.23,mirror},yshift=0pt]
-      (5,4) -- (5,9.5) node (cstk) [black,midway,xshift=0.8cm] {};
-
-      \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
-            {\fill[black] circle(1.5pt);}}}}
-      \draw[capa] (8.5,6.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\tiny{$(p,\gl)$}} (5,7.25);
-
-      \node[anchor=west] at (7.75,2) {$c_\var{next}$};
-      \draw[capa] (7.75,2) edge[-latex,postaction={decorate},out=180,in=0] node[right] {\tiny{$(p,\gl)$}}  (5.4,5.26);
-    \end{tikzpicture}
-    \caption{Visualization of ``$(\reg,\ms)$ is looking at $[i_0\dots i_n]$
-      followed by $c_{\var{next}}$''. $\reg(\pcreg)$ points to $i_0$ and
-      $c_{\var{next}}$ points to the address after $i_n$}
-  \label{fig:def-intuition-looking-at}
-  \end{subfigure}
-  \begin{subfigure}{.12\linewidth}
-$\phantom{te}$
-  \end{subfigure}
-  \begin{subfigure}{0.43\linewidth}
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-    % recurrent parts
-      \scope
-      \clip (-.1,-.1) rectangle (4.6,11.1);
-      \fill[fill=white] (0,0) rectangle (4.5,11);
-      \fill[color=gray!20] (0,9.5) rectangle (4.5,11);
-      \fill[color=gray!20] (0,0) rectangle (4.5,2);
-      \draw (0,0) -- (0,11);
-      \draw (4.5,0) -- (4.5,11);
-
-      \draw (0,9.5) -- (4.5,9.5);
-      \draw (0,4) -- (4.5,4);
-      \draw (0,2) -- (4.5,2);
-      \endscope
-      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
-      \draw (-1.5,0) node {};
-      \draw (-1.5,15) node {};
-
-      \node at (8.5,7.5) { $\reg$ };
-      \begin{scope}
-        \clip (7.9,4.5) rectangle (9.1,7.1);
-        \foreach \x in {4,5,...,6}
-        {
-          \draw (8,\x) rectangle (9,\x + 1);
-        };
-      \end{scope}
-      \node[anchor=west] at (9.125,5.5) {\footnotesize\texttt{r\_stk}};
-
-      \node[rotate=0] at (2.25,7) {\Large{$\ms_{\stk,\var{unused}}$}};
-      \node[rotate=0] at (2.25,3) {\Large{$\ms_{\stk}$}};
-
-      \draw [decorate,decoration={brace,amplitude=5pt,raise=0pt,aspect=0.23,mirror},yshift=0pt]
-      (4.6,2) -- (4.6,9.5) node (cstk) [black,midway,xshift=0.8cm] {};
-
-      \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
-            {\fill[black] circle(1.5pt);}}}}
-      \draw[capa] (8.5,5.5) edge[-latex,postaction={decorate},out=200,in=10,densely dashed] node[above,sloped] {\tiny $\rwlx$} (5,3.75);
-
-    \end{tikzpicture}
-    \caption{Visualization of ``$\reg$ points to stack with $\ms_\stk$ used and
-      $\ms_{\stk,\var{unused}}$ unused''. $\reg(r_{\var{stk}})$ is a local
-      $\rwlx$-capability that points to the
-      top most address of $\ms_\stk$.}
-  \label{fig:def-intuition-points-to-stack}
-  \end{subfigure}
-
-  \caption{}
-  \label{fig:def-intuition}
-\end{figure}
-
-The \texttt{malloc} macro is also common in our examples, so we prove Lemma~\ref{lem:malloc-works} to help reason about it.
-The structure of the \texttt{malloc} lemma is close to that of the \texttt{scall} lemma.
-It also has requirements on the configuration just before \texttt{malloc} is executed (\ref{item:malloc:looking-at}-\ref{item:malloc:malloc-sat}) as well as requirements on the execution afterwards (\ref{item:malloc:hyp-cont}).
-It does not have any requirements on the callee as the \texttt{malloc} specification defines its behavior.
-\begin{lemma}[$\mathtt{malloc}$ works]
-  \label{lem:malloc-works}
-  If
-  \begin{enumerate}
-  \item \label{item:malloc:looking-at}\lookingat{(\reg,\ms)}{\mathtt{malloc}\;r\;k}{c_{\var{next}}}
-  \item $k \geq 0$
-  \item \linksto{(\reg,\ms)}{\malloc}{k}{c_\malloc}
-  \item $c_\malloc$ satisfies the $\malloc$ specification with $\iota_{\malloc,0}$
-  \item $W \futurestr [i \mapsto\iota_{\malloc,0}]$
-  \item $\memSat{\ms}{W}$
-  \item $\ms = \ms' \uplus \ms_{\var{footprint}}$
-  \item \label{item:malloc:malloc-sat} $\memSat{\ms_{\var{footprint}}}{[i \mapsto W(i)]}$
-  \item \label{item:malloc:hyp-cont} \emph{Hyp-Cont}
-    If
-    \begin{itemize}
-    \item $n' \leq n-1$
-    \item $\iota_\malloc \futurewk W(i)$
-    \item $\memSat[n']{\ms_{\var{footprint}}' \uplus \ms'}{W[i \mapsto \iota_\malloc]}$
-    \item $\memSat[n']{\ms_{\var{footprint}}'}{[i \mapsto \iota_\malloc]}$
-      \begin{equation*}
-        \reg'(r') = 
-        \begin{cases}
-          c_{\var{next}} & r' = \pcreg \\
-          ((\rwx,\glob),\start,\addrend,\addr) & r' = r \\
-          \reg(r) & r' \not\in \RegName_t \union \{\pcreg, r, r_1\}
-        \end{cases}
-      \end{equation*}
-    \item $\addrend - \start = k - 1$
-    \item $\dom(\ms_{\var{alloc}}) = [\start,\addrend]$
-    \item $\forall \addr \in [\start,\addrend]\ldotp \ms_{\var{alloc}}(\addr) = 0$
-    \end{itemize}
-    Then we have $\npair[n']{(\reg',\ms' \uplus \ms_{\var{footprint}}'  \uplus \ms_{\var{alloc}})} \in \observations(W[\iota_\malloc])$
-  \end{enumerate}
-  Then
-  \[
-    \npair{(\reg,\ms)} \in \observations(W)
-  \]
-\end{lemma}
-In the technical appendix~\citep{technical_appendix}, we also define a lemma for the macro used to create closures, \texttt{crtcls}.
-
-% The below remark seems unnecessary: isn't this true about any lemma in any proof?
-
-% We stress that these lemmas
-% are conveniences and are not strictly necessary to reason about programs for the
-% capability machine.
-In principle, one could reason solely in the semantic model
-% without the use of these lemmas, but that would lead to a lot of repetition.
-\end{toplas}
-
-\section{Examples}
-\label{sec:examples}
-\itoplassug{Suggestion: add a couple of short proof sketches.}
-In this section, we demonstrate how our formalization of capability safety allows us to prove local-state encapsulation and control-flow correctness properties for challenging program examples.
-The security measures of \sectionname~\ref{sec:stack-and-return-pointer} are deployed to ensure these properties.
-Since we are dealing with assembly language, there are many details to the formal treatment. Therefore, we omit some details in the lemma statements.
-The examples may look deceivingly short, but it is because we use the macro instructions described in \sectionname~\ref{sec:macros}.
-The examples would be unintelligible without the macros as each macro expands to multiple basic instructions.
-
-\subsection{Encapsulation of Local State}
-The programs \texttt{f1} and \texttt{f2} in \figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's encapsulation of local state.
-They are very similar: both store some local state, call an untrusted piece of code ($\adv$), and test whether the local state is unchanged after the call.
-They differ in the way they do this.
-Program \texttt{f1} uses our stack-based calling convention (captured by \texttt{scall}) to call the adversary, so it can use the available stack to store its local state.
-On the other hand, \texttt{f2} uses malloc to allocate memory for its local state and uses \texttt{call} a calling convention based on heap allocated activation records (described in Appendix~\ref{app:macros}) to invoke the adversarial code.
-
-\begin{figure}[t]
-  \centering
-
-  \begin{minipage}[t]{4.1cm}
-  \begin{lstlisting}
-f1: push 1
-    fetch $r_1$ $\adv$
-    scall $r_1$($[]$,$[]$)
-    pop $r_1$
-    assert $r_1$ 1
-    halt
-  \end{lstlisting}
-  \end{minipage}
-  \begin{minipage}[t]{4.1cm}
-  \begin{lstlisting}
-f2: malloc $r_l$ 1
-    store $r_l$ 1
-    fetch $r_1$ $\adv$
-    call $r_1$($[]$,$[r_l]$)
-    assert $r_l$ 1
-    halt
-  \end{lstlisting}
-  \end{minipage}
-  \caption{Two example programs that rely on local-state encapsulation. \texttt{f1} uses our stack-based calling convention. \texttt{f2} does not rely on a stack.}
-  \label{fig:prog-f1-and-f2}
-\end{figure}
-
-For both programs, we prove that if they are linked with an adversary,
-$\adv$, allowed to allocate memory but has no other capabilities, then
-the assertion will never fail during execution (see
-Lemmas~\ref{lem:correctness-f1} and~\ref{lem:correctness-f2} below).
-% \lars{shouldn't we say already
-% here that the adversary is a global capability and that is why,
-% intuitively, the local state is preserved ?} \lau{Due to the ``order
-% of control'', \texttt{f1} runs first. The lemma is quite restrictive
-% on the adversary in the sense that it only has access to malloc (and
-% whatever we give et access to). Specifically, it does not have access
-% to the stack pointer and because of that we don't need to check it
-% first. We could write that the lemmas are limited in their assumptions
-% to make them simpler and a more complicated example is coming up.}
-The two examples also illustrate the versatility of the logical relation.
-The logical relation is not specific to any calling convention, so we can use it to reason
-about both programs even though they use different calling conventions.
-
-% \begin{lemma}[Correctness lemma for \texttt{f1}]
-% \todo[inline]{The lemma copied directly from the TR. To be deleted later!}
-%   \label{lem:correctness-f1}
-%   Let
-%   \begin{align*}
-%     c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}+\olf) \\
-%     c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1}-\olf,\mathtt{1f},\mathtt{f1}) \\
-%     c_\malloc & \defeq ((\entry,\glob),\start_\malloc,\addrend_\malloc,\start_\malloc+\olf) \\
-%     c_{\var{stk}} & \defeq ((\rwlx,\local),\start_\stk,\addrend_\stk,\start_\stk-1) \\
-%     c_\link & \defeq ((\readonly,\glob),\link,\link+1,\link)\\
-%     \reg & \in \Regs \\
-%     m & \defeq \hs_{f1} \uplus 
-%         \hs_\flag \uplus                
-%         \ms_{\var{link}} \uplus 
-%         \hs_\adv \uplus 
-%         \ms_{\malloc} \uplus 
-%         \ms_{\var{stk}} \uplus
-%         \ms_{\var{frame}} 
-%   \end{align*}
-%   and
-%   \begin{itemize}
-%   \item $c_\malloc$ satisfies the specification for malloc and $\iota_{\malloc,0}$ is the region from the specification.
-%   \end{itemize}
-%   where 
-%   \begin{align*}
-%     &\dom(\hs_{f1}) = [\mathtt{f1}-\olf,\mathtt{1f}] \\
-%     &\dom(\hs_\flag) = [\flag,\flag] \\
-%     &\dom(\ms_\link) = [\link,\link+1]\\
-%     &\dom(\ms_\stk) = [\start_\stk, \addrend_\stk]\\
-%     &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
-%     &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]} \qquad \text{ for all $n \in \nats$}
-%   \end{align*}
-%   and
-%   \begin{itemize}
-%   \item $\ms_{f1}(\mathtt{f1}-\olf) = ((\readonly,\glob),\link,\link+1,\link)$, $\ms_{f1}(\mathtt{f1}-\olf+1) = ((\readwrite,\glob),\flag,\flag,\flag)$, the rest of $\hs_{f1}$ contains the code of $f1$.
-%   \item $\ms_\flag = [\flag \mapsto 0]$
-%   \item $\ms_{\var{link}} = [\var{link} \mapsto c_\malloc, \var{link} + 1 \mapsto c_\adv]$
-%   \item $\hs_\adv(\start_\adv) = c_\link$ and $\forall \addr \in [\start_\adv+1,\addrend]\ldotp \ms_\adv(a) \in \ints$
-%   \end{itemize}
-%   if 
-%   \[
-%     (\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step[n] (\halted,m'),
-%   \]
-%   then
-%   \[
-%     m'(\flag) = 0
-%   \]  
-% \end{lemma}
-
-In order to formulate results about \texttt{f1} and
-\texttt{f2}, we need a way to observe whether the assertion
-fails. To this end, we assume they have access to a flag (an address in memory).
-If the assertion fails, then the flag is set to $1$ and execution halts. 
-\begin{lemma}[\texttt{f1} is correct]
-  \label{lem:correctness-f1}
-  Let
-\[
-    \begin{array}{r@{}lr@{}l}
-    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
-    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
-    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
-    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
-    \end{array}
-\]
-where each of the capabilities have an appropriate range of authority and
-pointer\footnote{These assumptions are kept intentionally vague for brevity.
-  Full statements are in the Appendix.}.
-Furthermore
-  \begin{itemize}
-  \item $c_\malloc$ satisfies the specification for malloc with
-    $\iota_{\malloc,0}$
-  \item $\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc,0}]}$
-  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
-  \item $\ms_\flag(\flag) = 0$
-  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
-  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
-  \end{itemize}
-  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
-  then $m'(\flag) = 0$
-\end{lemma}
-
-To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start configuration is safe (in the $\observations$ relation) for a world with a permanent region that requires the assertion flag to be 0.
-By Lemma~\ref{lem:anti-red-obs}, it suffices to show that the configuration is safe after some reduction steps.
-We then use the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}) by which it suffices to show that (1) the configuration that \texttt{scall} will jump to is safe and (2) the configuration just after \texttt{scall} is done cleaning up is safe.
-We use the Fundamental Theorem to reason about the unknown adversarial code as described in \sectionname~\ref{sec:reasoning}, but notice that the adversary capability is an enter capability, which the Fundamental Theorem says nothing about.
-Luckily, the enter capability has $\exec$-permission after the jump at which point the Fundamental Theorem applies.
-
-We have a similar lemma for \texttt{f2}:
-\begin{lemma}[\texttt{f2} is correct]
-  \label{lem:correctness-f2}
-  Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} but assuming no stack pointer and assuming $c_{f2}$ points to \texttt{f2},
-  if $(\reg\update{\pcreg}{c_{f2}},m) \step^* (\halted,m')$, then $m'(\flag) = 0$.
-\end{lemma}
-
-\subsection{Well-Bracketed Control-Flow} 
-The stack-based calling convention \texttt{scall} ensures well-bracketed control-flow.
-This is illustrated by program examples \texttt{f3} and \texttt{g1} in \figurename~\ref{fig:prog-f3-and-g1}.
-
-The program \texttt{f3} has two calls to an adversary.
-In order for the assertion on line 5 to succeed, the calls must be well-bracketed.
-If the adversary were able to store the return pointer from the first call and invoke it in the second call, then \texttt{f3} would have $2$ on top of its stack and the assertion would fail.
-However, the security measures in \sectionname~\ref{sec:stack-and-return-pointer} prevent this attack.
-Specifically, the return pointer is local, so it can only be stored on the stack.
-However, the part of the stack that is accessible to the adversary is cleared before the second invocation preventing attempts to store the return pointer.
-In fact, the following lemma shows that there are also no other attacks that can break well-bracketedness of this example, i.e.\ the assertion never fails.
-It is similar to the two previous lemmas:
-
-\begin{comment}
-\newsavebox{\testit}
-\begin{lrbox}{\testit}
-  \begin{lstlisting}
-g1:  malloc $r_2$ 1
-     store $r_2$ 0
-     move $r_3$ $\pcreg$
-     lea $r_3$ $\var{off}$
-     crtcls $[(x, r_2)]$ $r_3$
-     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
-     jmp $r_0$
-f4:  reqglob $r_1$
-     prepstk $r_\stk$
-     store $x$ 0
-     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
-     store $x$ 1
-     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
-     load $r_1$ $x$
-     assert $r_1$ 1
-     mclear $r_\stk$
-     rclear $\RegName \setminus \{r_0,\pcreg\}$
-     jmp $r_0$
-\end{lstlisting}
-\end{lrbox}
-
-\newsavebox{\progfthree}
-\begin{lrbox}{\progfthree}
-  \begin{lstlisting}
-f3: push 1
-    fetch $r_1$ $\adv$
-    scall $r_1$($[]$,$[r_1]$)
-    pop $r_2$
-    assert $r_2$ 1
-    push 2
-    scall $r_1$($[]$,$[]$)
-    halt
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}[t]
-  \centering
-  \subfloat{\usebox{\testit}}
-  \subfloat{\usebox{\progfthree}}
-  \caption{test}
-\end{figure}
-\end{comment}
-\begin{figure}[t]
-  \centering
-  \begin{minipage}[t]{.25\linewidth}
-  \begin{lstlisting}
-f3: push 1
-    fetch $r_1$ $\adv$
-    scall $r_1$($[]$,$[r_1]$)
-    pop $r_2$
-    assert $r_2$ 1
-    push 2
-    scall $r_1$($[]$,$[]$)
-    halt
-\end{lstlisting}
-  \end{minipage}%
-  \begin{minipage}[t]{.37\linewidth}
-  \begin{lstlisting}
-g1:  malloc $r_2$ 1
-     store $r_2$ 0
-     move $\pcreg$ $r_3$
-     lea $r_3$ $\var{off}$
-     crtcls $[(x, r_2)]$ $r_3$
-     rclear $\RegName \setminus \{\pcreg,r_0,r_1 \}$
-     jmp $r_0$
-f4:  reqglob $r_1$
-     prepstk $r_\stk$
-     ${\text{\tiny\itshape(continues in next column)}}$
-\end{lstlisting}
-  \end{minipage}%
-  \begin{minipage}[t]{.37\linewidth}
-\begin{lstlisting}[firstnumber=last]
-     ${\text{\tiny\itshape(continued from previous column)}}$
-     store $x$ 0
-     scall $r_1$($[]$,$[r_0,r_1,r_{\var{env}}]$)
-     store $x$ 1
-     scall $r_1$($[]$,$[r_0,r_{\var{env}}]$)
-     load $r_1$ $x$
-     assert $r_1$ 1
-     mclear $r_\stk$
-     rclear $\RegName \setminus \{r_0,\pcreg\}$
-     jmp $r_0$
-\end{lstlisting}
-  \end{minipage}
-  \caption{ Two programs that rely on well-bracketedness of
-    \texttt{scall}s to function correctly. $\var{off}$ is the
-    offset to \texttt{f4}.}
-  \label{fig:prog-f3-and-g1}
-\end{figure}
-
-\begin{lemma}
-  \label{lem:correctness-f3}
-  Making similar assumptions about capabilities and linking as in
-  Lemma~\ref{lem:correctness-f1} and assuming $c_{f3}$ points to \texttt{f3}
-  if $(\reg\update{\pcreg}{c_{f3}}\update{r_\stk}{c_\stk},m) \step^*
-  (\halted,m')$, then $m'(\flag) = 0$.
-\end{lemma}
-
-The final example, \texttt{g1} with \texttt{f4}, is a faithful translation of a tricky example known from the literature (known as the awkward example)~\citep{pitts_operational_1998,Dreyer:jfp12}.
-For comparison, we show a version of the original example in Figure~\ref{fig:awkward}, highlighting the code locations corresponding to \texttt{g1} and \texttt{f4}.
-
-\begin{wrapfigure}{R}{.35\textwidth}
-  \centering
-  \begin{tabular}{c} % see https://tex.stackexchange.com/questions/5818/how-to-center-a-listing
-  \begin{lstlisting}[gobble=2,escapeinside=||,keywords={fun, let, in, ref, assert}]
-  fun _ => 
-  |\color{gray}{g1:}| let x = ref 0 in
-      fun adv =>
-  |\color{gray}{f4:}|   x := 0;
-        adv();
-        x := 1;
-        adv();
-        assert(x == 1)
-  \end{lstlisting}
-  \end{tabular}
-  \caption{The original awkward example from \citet{pitts_operational_1998,Dreyer:jfp12}, in ML notation.}
-  \label{fig:awkward}
-\end{wrapfigure}
-
-Let us first look at this ML program.
-At the top-level, it is a lambda function that can be invoked by the context.
-When it is invoked, it allocates a fresh mutable variable $x$ of integer type that initially contains the value $0$.
-Next, the function returns a second closure (let's call this $\textit{cl}_x$) that can be invoked by the context whenever it chooses.
-$\textit{cl}_x$ itself takes a callback function \texttt{adv} that it will invoke twice after setting $x$ to $0$ and $1$ respectively.
-After the second invocation of \texttt{adv}, $\textit{cl}_x$ will verify that $x$ is still set to $1$.
-
-The assertion on line 8 of Figure~\ref{fig:awkward} should never fail.
-From the code, this seems natural since \texttt{adv} does not have access to $x$.
-Therefore after the second invocation on line 7, $x$ should still be in the state it was before that invocation.
-However, \texttt{adv} could have access to $\textit{cl}_x$ and could invoke $\textit{cl}_x$ again within the second invocation of \texttt{adv}.
-Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
-If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
-Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
-In other words, if we are able to prove that the assertion will never fail, this attack, and other similar attacks, on well-bracketed control flow are adequately prevented.
-
-Our low-level version of the awkward example in Figure~\ref{fig:prog-f3-and-g1}, consists of two parts, \texttt{g1} and \texttt{f4}, corresponding to the code locations marked in Figure~\ref{fig:awkward}.
-The program \texttt{g1} corresponds to the top-level closure in the ML code.
-It is a closure generator that generates closures with a mutable variable $x$ set to $0$ in its environment and \texttt{f4} as the program (note that we omit some calling convention security measures because the stack is not used in the closure generator).
-
-The program \texttt{f4} expects one argument, the callback \texttt{adv}, in \texttt{r1}.
-\texttt{f4} sets $x$ to $0$ and invokes \texttt{adv}.
-When it returns, \texttt{f4} sets $x$ to $1$ and calls \texttt{adv} again.
-When it returns the second time, \texttt{f4} asserts $x$ is still $1$ and returns.
-
-This example is more complicated than the previous ones because it involves a closure invoked by the adversary and an adversary callback invoked by us.
-As explained in \sectionname~\ref{sec:stack-and-return-pointer}, we need to check that (1) the stack pointer that the closure receives from the adversary has write-local permission and (2) the adversary callback is global.
-
-The attack that we explained above is actually more natural at this low-level.
-The callback \texttt{adv} now gets a return capability.
-As explained, it could invoke $\mathit{cl}_x$ again, during its second invocation, with a second callback function $\texttt{adv}'$ for $\mathit{cl}_x$ to invoke.
-If \texttt{adv} had some way of passing its return capability to $\texttt{adv}'$ (by storing it on the stack or in the heap, or hide it in an unused register), then the assertion could be made to fail.
-However, our calling convention prevents any of this from happening as we prove in the following lemma.
-
-\begin{lemma}
-  \label{lem:correctness-g1}
-  Let
-\[
-    \begin{array}{r@{}lr@{}l}
-    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
-    \end{array}
-\]
-  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
-  Then if $
-  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
-  then $m'(\flag) = 0$.
-\end{lemma}
-\begin{toplas}
-  \begin{proof}[Proof sketch]
-    Define a world $W_1$ with the following regions
-    \begin{itemize}
-    \item A malloc region, $\iota_{\malloc,0}$.
-    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
-    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
-    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\adv}$
-    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
-      accepts $\ms_\flag$ and $\ms_{g1}$.
-    \end{itemize}
-    If we can show
-    \begin{equation}
-      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
-  \end{equation}
-  \footnote{We ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
-
-  As described in \sectionname~\ref{sec:reasoning}, we use the FTLR (Theorem~\ref{thm:ftlr}) to reason about unknown code, so we use it to reason about $c_\adv$.
-  With the standard region $\iota^{nwl,p}$ chosen for the adversary, it is easy to show that the $\readCond{}$ and $\writeCond{}$ holds for $c_\adv$ which gives us $c_\adv \in \stder(W)$ by the FTLR.
-  In order to get \eqref{eq:g1:main-O-res}, we need to show that (a) the memory satisfies the world and (b) the register file is in the $\stdrr(W)$.
-  We have defined the world, so there (almost) is a one-to-one correspondence between regions in the world and memory segments in the initial configuration, so (a) easily follows.
-  In order to show (b), we need to show $c_\stk \in \stdvr(W)$ and $c_{g1} \in \stdvr(W)$.
-  The former follows easily from the choice of a $\iota^{pwl}$-region for the stack in $W$.
-  In order to argue the latter, we basically have to argue that $c_{g1}$, the closure generator, respects the world $W$.
-  This amounts to showing that the closures generated by \texttt{g1} also respect the invariants of $W$.
-  We reason about the local variable $x$ in the closure in the same way \citet{Dreyer:jfp12} does.
-  We ignore $x$ in the remainder of this proof sketch to focus on the parts of the proof related to the setting of the capability machine.
-
-  The capability for the generated closure is a global enter capability that we call $c_{f4}$.
-  The remainder of the proof amounts to showing that it is safe to return $c_{f4}$ to the adversary, i.e.
-  $c_{f4} \in \stdvr(W')$ where $W'$ is $W$ with relevant regions added after executing \texttt{g1}.
-  The adversary can use $c_{f4}$ whenever, so all we may assume about the configuration that $c_{f4}$ is invoked in is that the register-file $\reg$ and memory $\ms$ satisfies a world $W_1$ that is a private future world of $W'$, i.e.\ $\reg \in \stdrr(W_1)$ and $\memSat[]{\ms}{W_1}$.
-  We have to show that the invocation respects $W_1$ which corresponds to showing $(\reg[\pcreg \mapsto \updatePcPerm{c_{f4}}],\ms) \in \observations(W_1)$.
-  When $c_{f4}$ is invoked, we know exactly what instructions are executed up until the \texttt{scall}, namely it is ensured that the callback is global and the stack pointer has read/write-local/execute permission, and $x$ is set to 0.
-  Because we know part of the execution, we can apply the anti-reduction lemma (Lemma~\ref{lem:anti-red-obs}).
-
-  The next part in the execution is an \texttt{scall}, so, according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
-  For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
-  Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{cf.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
-  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.
-  In $W_2$, two new regions govern the stack.
-  One of the two new regions governs the caller stack frame, i.e.\ the part of the stack that contains the contents of private registers and stack recovery code.
-  The region that takes care of this is a temporary static region which ensures that our local stack frame remains unchanged during the execution of the callback.
-  The other new region takes care of the unused part of the stack which we are going to let the callback use for its execution.
-  This is taken care of by a standard region $\iota^{pwl}$-region which allows the callback to store any safe value on the stack.
-  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we assume that the arguments for the invocation of $c_{f4}$ are safe).
-  Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
-  Note also that arguing that the memory satisfies $W_2$ when we invoke the callback only works because we cleared the stack entirely (including the unused part) before the invocation.
-  Otherwise, it might contain local values that are only known to be safe in $W_1$, but for which we do not know that they will remain safe in the private future world $W_2$.
-  This part corresponds to arguing about Hyp-Callee in the \texttt{scall} lemma.
-
-  We need to argue that it is safe to give the return pointer that \texttt{scall} constructs to the adversary which corresponds to reasoning about the remainder of the execution of \texttt{f4} (corresponding to the Hyp-Cont case of the \texttt{scall} lemma).
-  The return pointer is a local capability, so we may assume that it is invoked in some configuration that satisfies $W_3 \futurewk W_2$.
-  This means that none of the temporary regions have been revoked, so the regions we replaced the old stack region with are still present in $W_3$.
-  We know exactly how the execution proceeds upon return (the recovery code is executed, and $x$ is set to 1).
-  We use the anti-reduction lemma to reason about the execution until the second \texttt{scall}.
-
-  For the \texttt{scall} itself we apply the \texttt{scall} lemma.
-  As the private stack has changed, we need to replace the two regions that govern the stack.
-  This means that the second invocation of the callback takes place in a world $W_4 \futurestr W_3$.
-  The reasoning about \texttt{scall} is similar to the first callback invocation.
-  The callback capability is still safe because it is global which basically covers the Hyp-Callee case.
-  For the Hyp-Cont case, we get a world $W_5 \futurewk W_4$ in which we need to argue that the remainder of the execution is safe.
-  At this point, we can use the anti-reduction lemma one final time to get to the point where we jump to the return pointer.
-
-  If we can argue that it is safe to jump to the return pointer that we got initially from the adversary, then the proof is done.
-  We have made no checks on the return pointer, so we have no idea whether it is local or global\footnote{We may assume that it is a capability that is executable when jumped to since otherwise the execution fails which is considered safe.}.
-  W.l.o.g.\ assume the return pointer is local.
-  This means that it is safe in public future worlds of $W_1$ (remember that it came from $\reg$ and $\reg \in \stdrr(W_1)$).
-  Hence we need to construct a world $W_6$ with all the temporary and revoked regions of $W_1$ (this corresponds to restoring the invariants the adversary relies on for its safe execution).
-  Further, for $W_6$ to be a public future world, for all these regions, $W_6$ must use the same region names as $W_1$. 
-  As we used the anti-reduction lemma to get to this point, $W_6$ must also be constructed to be a private future world of $W_5$.
-  Before we construct $W_6$, let us consider where each of the worlds we have seen so far came from: As illustrated in \figurename~\ref{fig:worlds-in-pf-g1}, we constructed $W_2$ and $W_4$.
-  These are the only private future worlds we have considered so far.
-  This means that we know exactly what changes they made.
-  In particular, we did not mask any of the temporary or revoked regions in $W_1$ with a permanent region.
-  This means that in $W_6$, we can reinstate every temporary or revoked region of $W_1$.
-  In the private future worlds we constructed, we only took transitions that are
-  public relative to $W_1$.
-  The worlds we were given, $W_3$ and $W_5$, may have taken arbitrary public transitions, but this is no problem with respect to the public future world relation.
-  In $W_5$ there were regions to handle the stack.
-  These regions need to be revoked as the $W_1$ stack regions replace them.
-  The non-stack regions in $W_5$ that are not present in $W_1$ are also added in $W_6$, which the public future world relation permits as it is extensional.
-  All in all, it is possible to construct $W_6$, so it is both a public future world of $W_1$ and a private world of $W_5$ which means that it is indeed safe to return to the adversary.
-
-  For the sake of presentation, we have omitted many details and made several simplifications in the above proof.
-  The complete proof can be found in the technical appendix~\citep{technical_appendix}.
-  \end{proof}
-\end{toplas}
-% As explained in \sectionname~\ref{sec:stack-and-return-pointer}, the
-% macro-instruction \texttt{\footnotesize{reqglob $r_1$}} checks that the callback
-% is global, essentially to make sure it is not allocated on the stack where it
-% might contain old stack pointers or return pointers. Otherwise, the
-% encapsulation of our local stack frame could be broken. In the proof of
-% Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
-% callback in a world that is only a private future world of the one where we
-% received the callback, precisely because we have invalidated the adversary's
-% local state (particularly their old stack and return capabilities). The callback
-% is still valid in this private future world, but only because we know that it is
-% global.
-
-% \itoplassug{Suggestion: Add an actual proof sketch}
-% In Lemma~\ref{lem:correctness-g1} the order of control has been
-% inverted compared to the previous lemmas. In this lemma, the adversary
-% assumes control first with a capability for the closure creator
-% \texttt{\footnotesize{g1}}. Consequently, we need to check that all
-% arguments are safe to use and that we clean up before returning in the
-% end. The inversion of control poses an interesting challenge when it
-% comes to reasoning about the adversary's local state during the
-% execution of \texttt{\footnotesize{f4}} and the callbacks where the
-% adversary should not rely on the local state from before the call of
-% \texttt{\footnotesize{f4}}. This is easily done by revoking all the
-% temporary regions of the world given at the start of
-% \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}}
-% returns, the adversary is again allowed to rely on its old local state
-% so we need to guarantee that the local state is unchanged. This is
-% important because the return pointer that \texttt{\footnotesize{f4}}
-% receives may be local, and the adversary is allowed
-% to allocate the activation record on the stack (just like we do) so
-% they can store and recover their old stack pointer after
-% \texttt{\footnotesize{f4}} returns.
-
-% \begin{toplas}
-% In \figurename~\ref{fig:worlds-in-pf-g1}, we illustrate how we accommodate this in
-% the proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds
-% for all the situations where control is passed to the adversary. The potentially
-% local return pointer received by \texttt{\footnotesize{f4}} from the adversary
-% is safe in $W_1$, so it can only be used in public future worlds. Let us see how
-% $W_6$ is constructed as a public future world of $W_1$, as well as a private
-% future world of $W_5$. The worlds $W_2$ and $W_4$ are constructed by revoking
-% all temporary regions and adding a $\iota^\pwl$ region for the stack we pass
-% away and static temporary regions (a region that accepts exactly one memory) for
-% our local stack as well as the adversary's local stack. The given worlds $W_3$
-% and $W_5$ are public future worlds in which the adversary chooses to return to
-% us. None of these worlds can have masked the temporary regions of $W_1$ with
-% permanent regions. $W_6$ is constructed by revoking the temporary regions of
-% $W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$ a public
-% future world of $W_1$. It is therefore safe to use the return pointer to return
-% to the adversary since we have restored validity of any local state they might
-% have stashed away.
-% \end{toplas}
-
-\begin{figure}[t]
-  \centering
-\begin{tikzpicture}[main node/.style={}, node distance=1.5cm]
-  \node[main node] (1) {$W_1$};
-  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f4} called}}};
-  \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};
-
-  \node[main node] (2) [above right of=1]{$W_2$};
-  \node[main node] (label2) [above of=2,yshift=-1cm] {\textit{\footnotesize{first callback}}};
-
-  \node[main node,left] (const) [above of=given,yshift=-0.4cm] {\parbox{1.5cm}{\footnotesize{Constructed:}}};
-
-  \node[main node] (3) [below right of=2]{$W_3$};
-  \node[main node] (label3) [below of=3,yshift=0.75cm] {\footnotesize{\textit{callback returns}}};
-
-  \node[main node] (4) [above right of=3]{$W_4$};
-  \node[main node] (label4) [above of=4,yshift=-0.75cm] {\footnotesize{\textit{second callback}}};
-
-  \node[main node] (5) [below right of=4]{$W_5$};
-  \node[main node] (label5) [below of=5,yshift=1cm] {\footnotesize{\textit{callback returns}}};
-
-  \node[main node] (6) [above right of=5]{$W_6$};
-  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f4} returns}}};
-  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
-       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
-       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
-       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
-       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
-\end{tikzpicture}
-\caption{Illustration of the worlds in the proof of
-  Lemma~\ref{lem:correctness-g1}. In the proof, the top row of worlds are
-  constructed by us, while the bottom row of worlds are given. $W_6$ is constructed such
-  that $W_6 \futurestr W_5$ and $W_6 \futurewk W_1$.}
-  \label{fig:worlds-in-pf-g1}
-\end{figure}
-
-\begin{comment}
-\begin{itemize}
-\item Ticket dispenser
-\item The awkward example and variants
-\item A sandboxing example?
-
-For example, an untrusted advertisement scenario with initialization code
-that registers a redraw callback. The redraw callback gets temporary
-read/write access to a framebuffer.
-
-\item Some compartmentalization result?
-\end{itemize}
-\end{comment}
-
-\section{Discussion}
-\label{sec:discussion}
-\itoplassug{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
-\itoplassug{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
-% p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
-\itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
-% * Also, the local capabilities seem related to Boyland's "Borrowing"
-%   e.g. that is now supported in Rust amongst other research languages
-%   (Pony etc) also mentioned in the ECOOP paper.
-% * the combination of E and L capabilities seems very interesting
-%   indeed.  I wonder if e.g. they could be used to support strong
-%   models of encapsulation e.g. like ownership types. (see eg LNCS
-%   7850)
-\noindent\textbf{Calling convention}\\
-\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
-Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
-  Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
-In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
-In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.
-The idea of their definition is to define a variant of a simple capability machine with a built-in stack as well as call and return instructions.
-The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
-They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
-Their full abstraction proof uses a logical relation, but it differs in a number of ways from the one we present.
-Their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
-This means that it could not be used to prove program correctness results similar to ours.
-Their calling convention is based on something called linear capabilities rather than local capabilities.
-Linear capabilities offer a different kind of limited revocation.
-\itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
-
-\emph{Performance and the requirement for stack clearing} The additional security measures of the calling convention described in \sectionname~\ref{sec:stack-and-return-pointer} impose an overhead compared to a calling convention without security guarantees.
-However, most require only a few atomic checks or register clearings on boundary crossings between trusted code and adversary, which should produce an acceptable performance overhead.
-The only exception is the stack clearing requirement that we have in two situations: when returning to the adversary and when invoking an adversary callback.
-As we have explained, we need to clear all of the stack that we are not using ourselves not just the part that we have actually used.
-In other words on every boundary cross between trusted code and adversary code, a potentially large region of memory must be cleared.
-
-First, contrary to what we explained before, we actually suspect that this overhead can be avoided when returning to the adversary.
-In that case, we now think it would suffice to clear only the (much smaller) part of the stack used by the trusted code itself.
-To understand this, it is useful to take another look at the illustrations in Figure~\ref{fig:ret-adv2} related to this case.
-If we do not clear the stack upon return (as illustrated in Figure~\ref{fig:illu-clear-all}), then the adversary might have used that stack to store local capabilities they received in a previous invocation (see Figure~\ref{fig:illu-adv-data}).
-In other words, the stack clearing is necessary for revoking those local capabilities: the stack pointer and return pointer for the invocation illustrated in Figure~\ref{fig:illu-adv-data}.
-While this approach is safe, we now suspect that we could do without the revocation in this case.
-The stack pointer which the adversary was given access to in the stack frame depicted in Figure~\ref{fig:illu-call-adv} only carries authority that the adversary has access to in the higher stack frame anyway.
-Similarly, the return pointer is merely an entry pointer pointing into the trusted code's stack frame, and this is also a capability that the adversary we return to could have constructed themselves from their stack pointer.
-
-This improvement of our calling convention is a recent insight and not yet reflected in our examples and our proofs.
-We do believe the proofs could be updated to accomodate for this change, but it would require some non-trivial changes.
-Consider the awkward example proof, for example, we would have to ensure that the world $W_6$, depicted in Figure~\ref{fig:worlds-in-pf-g1} would not just be a private future world of $W_5$, but a public one (in addition to being a public future world of $W_1$).
-This would allow us to do without the stack clearing, but it would entail some changes to the regions we use, and an extra argument that the old adversary's stack and return pointer remain valid after clearing the trusted code's stack frame and relaxing the invariant that used to ensure it could not be modified.
-While this change is a clear improvement, we do not actually believe that it fundamentally changes the efficiency characteristics of the approach: the cost is halved, but remains asymptotically the same.
-
-A second important remark we want to make is that the need for stack clearing in our calling convention is an instance of a general caveat when using CHERI's local capabilities as a restricted form of capability revocation.
-Consider how our use of local capabilities can be interpreted as temporarily delegating the stack and return capability to callees and then revoking the granted authority after the callee returns.
-From this perspective, local capabilities are a general feature enabling this temporary delegation of authority for the duration of an invocation and this is also how they are described by the authors~\citep{Watson2015Cheri}.
-However, our requirement for stack clearing on boundary crossings is also general.
-Revoking authority that was granted temporarily using local capabilities requires clearing all memory for which the invokee had write-local authority (or at least erasing all local capabilities from that memory).
-Without micro-architectural support for efficiently clearing large ranges of memory, local capabilities can only be used for revocation in scenarios where the duration of a revocation is unimportant or the adversary only has write-local access to small amounts of memory.
-
-CheriBSD's use of local capabilities in CCall does not actually involve a form of revocation.
-CheriBSD's model involves a trusted stack manager that gives every compartment access to its own private stack using a local, write-local capability \citep{Watson2015Cheri}.
-The locality of the stack capability allows the trusted stack manager to prevent compartments from leaking their stack pointer in a boundary crossing, but those capabilities are never actually revoked.
-In fact, a compartment can easily store away such local capabilities in its own private stack and recover them there during future invocations.
-
-Since local capabilities seem intended to provide a restricted form of revocation, perhaps capability machines like CHERI should consider to provide special support for this requirement.
-Ideally, such support would take the form of a highly-optimized instruction for erasing a large block of memory.
-Recent work suggests that such a feature could perhaps be added to processors like CHERI, using a special hardware cache that tracks whether or not a memory location contains zero \citep{Joannou2017EfficientTaggedMemory}.
-
-% http://scale.eecs.berkeley.edu/papers/mmp-asplos2002.pdf
-% https://www.usenix.org/legacy/event/osdi08/tech/full_papers/zeldovich/zeldovich_html/index.html
-% http://users.ece.utexas.edu/~tiwari/pubs/MICRO-08-rangecache.pdf
-
-\emph{Modularity} It is important that our calling convention is modular, i.e.\ we do not assume that our code is specially privileged w.r.t.\ the adversary, and they can apply the same measures to protect themselves from us as we do to protect ourselves from them.
-More concretely, the requirements we have on callbacks and return pointers received from the adversary are also satisfied by callbacks and return pointers that we pass to them.
-For example, our return pointers are local capabilities because they must point to memory where we can store the old stack pointer, but the adversary's return pointers are also allowed to be local.
-Adversary callbacks are required to be global, but the callbacks we construct are allocated on the heap and also global.
-
-\emph{Arguments and local capabilities} 
-%Stack and return pointer
-%Local capabilities cannot be "freely used"
-%Safety dictated by LR
-%Need to dynamical check whether we overwrite our local stuff.
-Local capabilities are a central part of the calling convention as they are used
-to construct stack and return pointers. The use of local capabilities for the
-calling convention unfortunately limits the extent to which local capabilities
-can be used for other things. Say we are using the calling convention and
-receive a local capability other than the stack and return pointer, then we need
-to be careful if we want to use it because it may be an alias to the stack
-pointer. That is, if we first push something to the stack and then write to the
-local capability, then we may be (tricked into) overwriting our own local state.
-The logical relation helps by telling us what we need to ascertain or check in
-such scenarios to guarantee safety and preserve our invariants, but such checks
-may be costly and it is not clear to us whether there are practical scenarios
-where this might be realistic.
-
-%Pass capability on
-%Stack example
-%Pass argument by reference
-%Sofisticated deep safety analysis
-We also need to be careful when we receive a capability from an adversary that
-we want to pass on to a different (instance of the) adversary. It turns out that
-the logical relation again tells us when this is safe. Namely, the logical
-relation says that we can only pass on arguments that are safe in the world we
-invoke the adversary in. For instance, when we receive a stack pointer from an
-adversary, then we may at some point want to pass on part of this stack pointer
-to, say, a callback. In order to do so, we need to make sure the stack pointer
-is safe which means that, if we have revoked temporary invariants, the stack
-must not directly or indirectly allow access to local values that we cannot
-guarantee safety of. When received from an adversary, we have to consider the
-contents of the stack unsafe, so before we pass it on, we have to clear it, or
-perform a dynamic safety analysis of the stack contents and anything it points
-to. Clearing everything is not always desirable and a dynamic safety analysis is
-hard to get right and potentially expensive.
-
-% Can local capabilities be used for other things?/Local capabilities only for stack?
-%Not unheard of - other machine also have things specific for language features.
-In summary, the use of local capabilities for other things than stack and return
-pointers is likely only possible in very specific scenarios when using our
-calling convention. While this is unfortunate, it is not unheard of that
-processors have built-in constructs that are exclusively used for handling
-control flow, such as, for example, the call and return instructions that exist
-in some instruction sets.
-% Old version of the above:
-% Carefully constructing stack and return
-% pointer plays a big part in the calling convention, but it is completely
-% undermined if an unsafe argument, say the full stack pointer that include the
-% caller's local stack frame, is passed as an argument of the call. The logical
-% relation dictates when it is safe to pass an argument to an adversary, namely it
-% must be a safe value. In order to illustrate what this means for local
-% capabilities, consider the stack capability. When the stack capability is
-% received from an adversary, then the contents of the stack is unknown and must
-% thus be considered unsafe, so before the stack capability can be passed on, we
-% must make sure that it only gives access to safe values. In order to meet this
-% requirement, we take a conservative approcah and simply clear everything on the
-% stack. Integers are always safe, so the stack capability can now safely be
-% passed on. While this approach works for the stack, it may in practice be a bit
-% too naive for other local capabilities received form an adversary. Say we
-% receive a capability from an adversary that happens to be the first part of the
-% stack. We want to pass this capability on, so we naively clear the contents it
-% gives access to which in the worst case means we overwrite our local data or the
-% activation record. While this is strictly speaking not unsafe (integers are
-% always) safe, it is not desirable in a real system. We would in principle also
-% like to be able to pass on arguments by reference, e.g a local capability for
-% some array of values. In this case, it would not be satisfactory to just clear the
-% entire thing. However, the capability still needs to be safe, so there would
-% need to be some kind of deep dynamic check that ensures that everything the
-% capability gives access to is indeed safe. In a real setting, this may again be
-% undesirable.
-% This also points to a more
-% general problem: When can we use local capabilities we have received from an
-% adversary. We at least need to do some kind of dynamic range check to make sure
-% that we do not inadverdently overwrite our local stack frame.
-
-\itoplassug{We could also include a brief discussion of concurrency and the fact that this probably wouldn't scale to it (but well-bracktedness is not the same in that case anyway.) }
-\itoplassug{Reviewer C, esop, wants to know what happens when there are several stack blocks (see tex for comment).
-We could here emphasize that while we talk about a single stack there could be multiple local read/write-local/execute-capabilities.
-If we get the stack from an untrusted source, which is the case in the awkward example, then we will use anything that \emph{looks like} the stack as the stack.
-However, due to the operational semantics for local capabilities all other ``stack blocks'' would have to be stored on the stack or in the register file.
-As the calling convention dictates that we clear these, the only place the alternative stack blocks will not be cleared is on a part of the stack that we do not have access to.
-The return capability would basically have to encapsulate this part of memory (perhaps indirectly) which means that from the moment we have been called, we will make sure there is only one stack available until our program returns from the initial call to it.}
-%p.10 "to prevent this from happening, we need to make sure the stack capability carries RWLX authority since the system wide assumption then tells us that the adversary cannot have global capabilities to our local stack".
-It feels like the "L" capabitility is being overloaded to mean not just "this memory block has write-local authority and can contain local capabilities", but also to mean "this memory block is my stack and no other block is".
-What happens when there are several stack blocks, e.g.
-one per thread?
-\emph{Single stack} A single stack is a good choice for the simple capability machine presented here because it works well with higher-order functions.
-An alternative to a single stack would be to have a separate stack per component.
-The trouble with this approach is that, with multiple stacks and local stack pointers, it is not clear how components would retrieve their stack pointer upon invocation without compromising safety.
-A safe approach could be to have stack pointers stored by a central, trusted stack management component, but it is not clear how that could scale to large numbers of separate components.
-Handling large numbers of components is a requirement if we want to use capability machines to enforce encapsulation of, for example, every object in an object-oriented program or every closure in a functional program.
-\\\\
-\begin{toplas}
-  \noindent \textbf{Capability machine formalization}\\
-  \emph{Simplifications} Our capability machine formalization assumes unbounded integers and an infinite address space.
-  Further, it has a much simpler instruction set than that of a full ISA.
-  By making these simplifications, we avoid tedious details but end up with a less realistic machine.
-  However, the intend of this work is to gain ground in the formal work necessary to prove properties about low-level assembly programs on a capability machine not to apply it to a full fledged capability machine.
-
-  We do not believe that any of our simplifications are beyond reason and expanding the result seems plausible.
-  If we added unbounded integers, we would have to change the operational
-  semantics to take over and under flows into account. This could be achieved
-  by changing the \texttt{plus} and \texttt{minus} instruction to use modulo
-  arithmetic.
-  If we added a bounded address space, then we would have to take memory out of
-  bounds into account. This would require changes to the memory
-  operations (\texttt{store} and \texttt{load}) and the step relation of the operational semantics.
-  The malloc specification would also need to be updated as it would have to
-  signal failure when it runs out of memory.
-  Finally, expanding the instruction set to a realistic ISA would add a handful
-  of instructions that we would need to reason about in the proofs.
-  All in all, it would add a lot of tedious work to expand this result to a
-  realistic machine. The amount of details in the proofs is already at the
-  threshold for what should be done with pen and paper, so expanding this work
-  to a realistic machine would require mechanized proofs that can take care of
-  the tedious details.
-\end{toplas}
-\\\\
-\begin{toplas}
-\noindent \textbf{Reasoning about capability machine programs}\\
-\emph{Limitations} The logical relation defined in
-\sectionname~\ref{sec:logical-relation} allows us to reason about capability machine
-programs. A limitation w.r.t. previous work is that the logical relation is
-tailored towards proofs of a specific class of properties.
-
-Imagine, for example, that we invoke a block of adversary code in such a way
-that it only ever receives capabilities within a specific range of memory. After
-the code returns, we may try to prove that any capabilities passed back to us in
-the registers are still confined to that range of memory. The property talks
-about a the specific implementation of a higher-order value rather than its behavior,
-like the invariants that are required/preserved when we use it.
-
-Such properties are hard to prove in our model.
-For the example, it would be easy to conclude that the returned values are in the value relation (see \figurename~\ref{fig:logrel}).
-This gives us a lot of behavioral information, like conditions under which the capabilities are safe to use and invariants that will be preserved when we do, but it does not tell us much about the range of authority of the capability.
-As a very concrete example, capabilities with permission $\noperm$ are always in the value relation irrespective of their range of authority.
-Behaviorally, this makes perfect sense; $\noperm$-capabilities cannot be used
-for anything.
-
-For our purposes, this restriction is unproblematic since we are only interested in proving behavioral properties (e.g., an assertion will never fail).
-In other situations, however, we may be interested in proving properties like the ones that are often considered in object capability literature: confinement, no authority amplification etc.
-Although such properties are more restrictive and tough to use for reasoning, \citet{Devriese:2016ObjCap} have demonstrated how a logical relation like ours can be adapted to also support them by quantifying the logical relation over a custom interpretation of effectful computations and the type of references.
-We expect their solution can be readily adapted to our setting modulo some details (like the fact that we do not just have read-write capabilities but also others).
-\end{toplas}
-% % \emph{No program logic} TODO: discuss this? 
-\\\\
-\noindent\textbf{Logical relation}\\
-\emph{Single orthogonal closure} The definitions of $\stder$ and $\stdvr$ in \figurename~\ref{fig:logrel} apply a single orthogonal closure, a new variant of an existing pattern called biorthogonality.
-Biorthogonality is a pattern for defining logical relations~\citep{krivine_classical_1994,pitts_operational_1998} in terms of an observation relation of safe configurations (like we do).
-The idea is to define safe evaluation contexts as the set of contexts that produce safe observations when plugged with safe values and define safe terms as the set of terms that can be plugged into safe evaluation contexts to produce safe observations.
-This is an alternative to more direct definitions where safe terms are defined as terms that evaluate to safe values.
-An advantage of biorthogonality is that it scales better to languages with control effects like call/cc.
-Our definitions can be seen as a variant of biorthogonality; we take only a single orthogonal closure.
-We do not define safe evaluation contexts but immediately define safe terms as those that produce safe observations when plugged with safe values.
-This is natural because we model arbitrary assembly code that does not necessarily respect a particular calling convention.
-Return pointers are in principle values like all others and there is no reason to treat them specially in the logical relation.
-
-Interestingly, \citet{Hur:2011:KLR:1926385.1926402} also use a step-indexed,
-Kripke logical relation for an assembly language (for reasoning about correct
-compilation from ML to assembly). However, because they only model non-adversarial
-code that treats return pointers according to a particular calling convention,
-they can use standard biorthogonality rather than a single orthogonal closure
-like us.
-
-\emph{Public/private future worlds} A novel aspect of our logical relation is
-how we model the temporary, revokable nature of local capabilities using public/private
-future worlds. The main insight is that this special nature generalizes that of the syntactically-enforced
-unstorable status of evaluation contexts in lambda calculi without control
-effects (of which well-bracketed control flow is a consequence). To reason about
-code that relies on this (particularly, the original awkward example),
-\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
-contexts using Kripke worlds with public and private future world relations.
-Essentially, they allow relatedness of evaluation contexts to be monotone with
-respect to a weaker future world relation (public) than relatedness of values,
-formalizing the idea that it is safe to make temporary internal state
-modifications (private world transitions, which invalidate the continuation,
-but not other values) while an expression is performing internal steps, as long
-as the code returns to a stable state (i.e. transitions to a public future world
-of the original) before returning. We
-generalize this idea to reason about local capabilities: validity of local
-capabilities is allowed to be monotone with respect to a weaker future-world
-relation than other values, which we can exploit to distinguish between state
-changes that are always safe (public future worlds) and changes that are only
-valid if we clear all local capabilities (private future worlds). Our future
-world relations are similar to DNB's (for example, our proof of
-the awkward example uses exactly the same state transition system),
-but they turn up in an entirely different place in the logical relation: rather
-than using public future worlds for the special syntactic category of evaluation
-contexts, they are used in the value relation depending on the
-locality of the capability at hand. Additionally, our worlds are a bit more
-complex because, to allow local memory capabilities and write-local
-capabilities, they can contain (revokable) temporary regions that are only
-monotonous w.r.t. public future worlds, while DNB's worlds are entirely
-permanent.
-
-\emph{Local capabilities in high-level languages} We point out that
-local capabilities are quite similar to a feature proposed for the high-level
-language Scala: \citet{osvald_gentrification_2016}'s second-class or local
-values. They are a kind of values that can be provided to other code for
-immediate use without allowing them to be stored in a closure or reference for
-later use. We believe reasoning about such values will require techniques similar 
-to what we provide for local capabilities.
-
-\begin{toplas}
-  \emph{Why use a logical relation and not a simpler proof technique?}
-  The concept of a capability exists on different levels of abstraction on computers, so capabilities have been studied and safety properties proven about them in other contexts than assembly languages.
-  Traditionally, logical relations have not been used for safety property proofs, so why do we need one here?
-  To answer this question let's compare this work to \citet{sewel_integrity_2011} where they prove \textit{integrity} and \textit{authority confinement} for the seL4 Microkernel.
-  % Describe what they prove
-  The security properties they show are relative to a security policy that specifies the upperbound of capabilities a subject in the system can posses.
-  The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject a can have a write capability for memory B.}
-  However, their policy are, in a sense, binary: subject a can either not write or write anything to memory B.
-  Our approach is much more fine-grained and allows us to define, for example, a policy that subject a can write to memory B, but only if the value is an even number.
-  Because our invariants are indexed by worlds themselves, we can even define a policy that capabilities written to memory B must respect certain other invariants themselves.
-  It is exactly the expressive invariants that enable us to prove the awkward example.
-  We place a protocol on the memory where variable \texttt{x} is stored that says that it sometimes can be either $0$ or $1$ and at other times it has to be $1$.
-
-  Capability safety is not just about the permission a capability carries, it is also about how the capability is used.
-  Say for instance, a program has a write capability that will only be used to write even numbers.
-  If this is the only write capability for that part of memory, then we should be able to rely on that part of memory only containing even numbers.
-  In order to rely on this in our proofs, our notion of capability safety must be able to take the meaning of the program into account and express the memory invariant.
-\end{toplas}
-\itoplassug{Reviewer C, popl, would like to know how local capabilities relate to borrowing (see tex comment).}
-
-\section{Related Work}
-\label{sec:related-work}
-In this section, we summarize how our work relates to previous work.
-We do not repeat the work we discussed in \sectionname~\ref{sec:discussion}.
-
-Capability machines originate with \citet{Dennis:1966:PSM:365230.365252} and we refer to \citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview of previous work.
-The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is a simple but representative model modeled mainly after the M-Machine~\citep{Carter:1994:HSF:195473.195579} (the enter pointers resemble the M-Machine's) and CHERI~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the memory and local capabilities resemble CHERI's).
-The latter is a recent and relatively mature capability machine. CHERI combines capabilities with a virtual memory approach in the interest of backwards compatibility and gradual adoption.
-As discussed, our local capabilities can cross module boundaries contrary to what is enforced by CHERI's default CCall implementation.
-
-Plenty of other papers\lau{If there are plenty, then I guess we should cite more than one?}
-enforce well-bracketed control flow at a low level but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
-This includes particularly the line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
-This line of work uses a quite different attacker model than us.
-They assume an attacker that is unable to execute code but can overwrite arbitrary data at any time during execution (to model buffer overflows).
-By checking the address of every indirect jump and using memory access control to prevent overwriting code, this work enforces what they call control-flow integrity formalized as the property that every jump will follow a legal path in the control-flow graph.
-As discussed in \sectionname~\ref{sec:discussion}, such a property ignores temporal properties and seems hard to use for reasoning.
-
-More closely related to our work are papers that use a trusted stack manager and
-some form of memory isolation to enforce control-flow correctness as part of a
-secure compilation
-result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}. Our work
-differs from theirs in that we use a different form of low-level security
-primitive (a capability machine with local capabilities rather than a machine
-with a primitive notion of compartments). Further, we do not use a trusted stack
-manager but a decentralized calling convention based on local
-capabilities. Also, both prove a secure compilation result from a high-level
-language which clearly implies a general form of control-flow correctness
-while we define a logical relation that can be used to reason about
-specific programs that rely on well-bracketed control flow.
-
-% \citet{skorstengaard_stktokens_2019} gives a definition of well-bracketed control flow based on overlay semantics.
-% Their definition characterizes the kind of well-bracketed control flow our examples rely on.
-% They define well-bracketed control flow for a capability machine by defining a new machine that is syntactically the same but adds a built-in call stack that is only accessed when a series of instructions corresponding to a call or return is executed.
-
-\lau{Reviewer C, esop, would like a better comparison with \citet{Devriese:2016ObjCap} (but also admits that they did not follow the below discussion).}
-Our logical relation is a unary, step-indexed Kripke logical relation with
-recursive worlds
-\citep{pitts_operational_1998,Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
-closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
-capability safety in a high-level JavaScript-like lambda calculus. Our
-Fundamental Theorem is similar to theirs and expresses capability safety of the
-capability machine. Because we are not interested in externally observable
-side-effects (like console output or memory access traces), we do not require
-their notion of effect parametricity. Our logical relation uses several ideas
-from previous work like Kripke worlds with regions containing state transition
-systems \citep{Ahmed:popl09}, public/private future worlds \citep{Dreyer:jfp12}
-(see \sectionname~\ref{sec:discussion} for a discussion), and biorthogonality
-\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.
-
-\citet{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
-of object capability patterns. The logic is based on
-Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
-separation logic, and is formalized in Coq building on the Iris Proof Mode for
-Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
-capability safety for a lambda-calculus (with concurrency) compared to the
-earlier work by \citet{Devriese:2016ObjCap}.
-% domi: rest of the paragraph could be cut?
-\begin{toplas}
-In the future, we would like to develop a new program logic for
-reasoning about capability safety for our capability machine model. In fact,
-we think the lemmas in Section~\ref{sec:reasoning} are suggestive of the style
-of results that could be captured in such a logic. We think Iris would also be
-a natural starting point for such an endeavour since Iris is a
-framework that can be instantiated with different programming languages. OCPL
-was able to leverage existing Iris specifications for a high-level programming language;
-for our capability machine model, however, it would be necessary to devise new
-kinds of specifications for our low-level programs with unstructured
-control-flow. It is likely that we could get inspiration from earlier work on
-logics for assembly programming languages, such as XCAP~\citep{xcap}.
-Building a logic around the semantic model presented here would remove some of the tedious repetitive proof details making it more realistic to prove properties about larger more realistic programs.
-
-If we want to scale this approach even further, we would like to reason at a higher level of abstraction namely at the level of a high-level language.
-That is, we would rather construct a program logic for a high-level programming language, so we can reason about the programs in the language we actually write them in and at the same time get guarantees about the compiled program.
-To achieve this, we would have to construct a secure compilation \citep{patrignani_secure_compilation_2019} that preserves the security abstraction of the high-level language.
-\end{toplas}
-\itoplassug{We could further add that this is not the level of abstraction that
-  one would like to reason. One would like to reason about the programs we
-  actually write - that is programs written in high-level languages. If we had a
-  fully-abstract compiler from a nice high-level language to this low-level
-  machine, then we could reason about the program in the high-level language and
-  then because the compiler is fully-abstract, we would retain all the nice
-  properties from the high-level language. How does this work fit in to this
-  picture? The calling convention ensures two properties that we expect our nice
-  high-level languages to have, namely well-bracketedness and local-state
-  encapsulation. If we want to do fully-abstract compilation, then we need some
-  way to guarantee this after compilation. Further, logical relations are often
-  used in full-abstractness proofs. This work expands our knowledge about
-  logical relations should be defined for low-languages and in particular
-  capability machines. (maybe also mention that enforcement necessary for
-  fully-abstract compilation. Could be capability, but it could also be
-  something else - we just need something which we do not have now.) }
-
-El-Korashy also defined a formal model of a capability machine, namely CHERI,
-and uses it to prove a compartmentalization
-result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
-correctness). He also adapts control-flow integrity (see above) to the machine
-and shows soundness, seemingly without relying on capabilities.
-
-\begin{acks}
-This research was supported in part by the ModuRes Sapere Aude Advanced Grant
-from The Danish Council for Independent Research for the Natural Sciences (FNU). 
-Support for an STSM was received from COST Action EUTypes (CA15123).
-Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
-\end{acks}
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
-
-%% Bibliography
-\bibliography{references}
-
 %% Appendix
@@
 \end{align*}
-
-\end{document}

### 2019-08-12 toplas/paper.tex
[TOPLAS] Typo

@@
 Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, \itoplas{creates a view with all capabilities that may be available at some point in time.
-By removing the time aspect, we may have to consider capabilities that are actually not available at a given point in time.}
+By removing the time aspect, we may have to consider capabilities that are actually not available at a given point in time.
   Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}

### 2019-08-12 toplas/paper.tex
[TOPLAS] Mention xi in paragraph about world existence
[TOPLAS] Add nsubset definition to appendix in paper

@@
 The formulation of the lemma is technical, so we recommend that non-expert readers ignore the technicalities and accept that there exists a set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying the (recursive) equations in the theorem (where the $\blater$ operator can be safely ignored\footnote{The interested reader can find a brief coverage of \cofe{}'s and $\blater$ in Appendix~\ref{app:cofe}.}).
+\begin{toplas}
+The solution $\Wor$ is some \cofe{} that we do not know much about that is, essentially, isomorphic to the our worlds.
+The isomorphism $\xi$ allows us to move between the solution and the worlds we can work with and respects the future world relations.
+\end{toplas}
+
 % Short technical explanation
@@
 \end{mathpar}
+where $\nsubeq$ is defined as follows: define erasure for step-indexed sets as
+\[
+  \lfloor A \rfloor_n = \{ \npair[m]{a} \in A \mid m < n \}
+\]
+and define $\nsubeq$ as
+\[
+  A \nsubseteq B \texttt{ iff } \lfloor {A} \rfloor_n \subseteq \lfloor {B} \rfloor_n
+\]
 %

### 2019-08-15 popl-2019-journal-v/proving-full-abstraction.tex
Minor typesetting improvements.
Typesetting fix.

@@
       \begin{array}{l}
-        n = 0 \vee{}\\
+        n = 0 \text{ or }\\
         x \nequal[n-1]_X x'
       \end{array}
-    \right.
+    \right. \qedhere
   \]
@@
   \]
-  to get a preordered c.o.f.e.\, we use subset as the preorder.
+  to get a preordered c.o.f.e.\ we use subset as the preorder.
 \end{definition}
@@
 
-Next, we define a c.o.f.e. of monotone, non-expansive functions
+Next, we define a c.o.f.e.\ of monotone, non-expansive functions
 \begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
@@
 \end{definition}
-In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\, and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
+In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\  and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
 

### 2019-08-15 toplas/paper.tex
fix typo

@@
   The security properties they show are relative to a security policy that specifies the upperbound of capabilities a subject in the system can posses.
-  The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject a can have a write capability for memory B.}
+  The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject A can have a write capability for memory B.}
   However, their policy are, in a sense, binary: subject a can either not write or write anything to memory B.

### 2019-08-19 toplas/paper.tex
[TOPLAS] Rewrite extensional discussion

@@
 \begin{toplas}
-  In order to be able to solve the recursive domain equation, we want the future world relations to be extensional.
   Intuitively, the future world relation specifies how memory protocols may change over time.
-  If the future world relation was not extensional, then the protocols could just disappear entirely regardless of what they promise.
-  A piece of memory can still be repurposed under the right circumstances, i.e., if the protocol allows it under the future world relation used.
-  In practice, the extensionality of future worlds means that a future world cannot forget about region names in a past world.
-  This is the reason we use masks rather than deleting the regions altogether.
-  It does not, however, matter what kind of region is used as a mask; it is the region name we must retain.
-  This is why we allow revoked regions to be replaced by other regions in the two future world relations.
-  This may seem unnecessary as the future world relation allows new regions to be added.
-  However, for technical reasons in private future worlds, we need to be able to reinstate old regions with the same region names as they had in a past world.
-  Further, we do not want public future worlds to be able to prevent reinstation in private future worlds.
-  Private future worlds cannot revoke permanent regions, so we cannot allow public future worlds to use permanent regions as masks for revoked regions.
+  Rather than deleting regions in future worlds, we follow the approach in \citet{Thamsborg:2011:KLR:2034773.2034831,Ahmed2004semantics} and use masks to signal which regions are active.
+  This approach implies that the future world relation is a preorder and hence we can use the
+  method in \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} to solve the recursive world equation.
 \end{toplas}

### 2019-08-19 toplas/paper.tex
[TOPLAS] Add sentence to Sewel discussion

@@
   The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject A can have a write capability for memory B.}
-  However, their policy are, in a sense, binary: subject a can either not write or write anything to memory B.
-  Our approach is much more fine-grained and allows us to define, for example, a policy that subject a can write to memory B, but only if the value is an even number.
+  However, their policy are, in a sense, binary: subject A is either allowed to write to memory B or not.
+  If writing is allowed, then any value can be written, as long as that value is itself legal.
+  \citet{sewel_integrity_2011}'s results are not parametrised by a (possibly more restrictive) policy that subject A should observe on memory B.
+  Our approach is much more fine-grained and allows us to define, for example, a policy that subject A can write to memory B, but only if the value is an even number.
   Because our invariants are indexed by worlds themselves, we can even define a policy that capabilities written to memory B must respect certain other invariants themselves.

### 2019-08-19 toplas/paper.tex
[TOPLAS] Update CFI discussion

@@
 Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, \itoplas{creates a view with all capabilities that may be available at some point in time.
-By removing the time aspect, we may have to consider capabilities that are actually not available at a given point in time.
-  Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
-In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
+  Hence control-flow-graph based formulations lead one to  consider capabilities that are actually not available at a given point in time.
+  Example~\texttt{g1} relies on a more fine-grained view of the control flow, in particular
+  when returning from the higher-order call backs.}
+In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than control-flow-graph based formulations, although this is not made very explicit.
 In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.

### 2019-08-20 toplas/paper.tex
typo and comment in TODO file

--- 2019-08-19 toplas/paper.tex
@@
 \begin{toplas}
+  In order to be able to solve the recursive domain equation, we want the future world relations to be extensional.
   Intuitively, the future world relation specifies how memory protocols may change over time.
-  Rather than deleting regions in future worlds, we follow the approach in \citet{Thamsborg:2011:KLR:2034773.2034831,Ahmed2004semantics} and use masks to signal which regions are active.
-  This approach implies that the future world relation is a preorder and hence we can use the
-  method in \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} to solve the recursive world equation.
+  If the future world relation was not extensional, then the protocols could just disappear entirely regardless of what they promise.
+  A piece of memory can still be repurposed under the right circumstances, i.e., if the protocol allows it under the future world relation used.
+  In practice, the extensionality of future worlds means that a future world cannot forget about region names in a past world.
+  This is the reason we use masks rather than deleting the regions altogether.
+  It does not, however, matter what kind of region is used as a mask; it is the region name we must retain.
+  This is why we allow revoked regions to be replaced by other regions in the two future world relations.
+  This may seem unnecessary as the future world relation allows new regions to be added.
+  However, for technical reasons in private future worlds, we need to be able to reinstate old regions with the same region names as they had in a past world.
+  Further, we do not want public future worlds to be able to prevent reinstation in private future worlds.
+  Private future worlds cannot revoke permanent regions, so we cannot allow public future worlds to use permanent regions as masks for revoked regions.
 \end{toplas}
@@
   The security policy is expressed as a collection of possible capabilities in the system, e.g.\ \textit{subject A can have a write capability for memory B.}
-  However, their policy are, in a sense, binary: subject A is either allowed to write to memory B or not.
-  If writing is allowed, then any value can be written, as long as that value is itself legal.
-  \citet{sewel_integrity_2011}'s results are not parametrised by a (possibly more restrictive) policy that subject A should observe on memory B.
+  However, their policies are, in a sense, binary: subject A can either not write or write anything to memory B.
   Our approach is much more fine-grained and allows us to define, for example, a policy that subject A can write to memory B, but only if the value is an even number.

### 2019-08-20 toplas/paper.tex
typo

--- 2019-08-19 toplas/paper.tex
@@
   Example~\texttt{g1} relies on a more fine-grained view of the control flow, in particular
-  when returning from the higher-order call backs.}
+  when returning from the higher-order callbacks.}
 In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than control-flow-graph based formulations, although this is not made very explicit.

### 2019-08-24 toplas/paper.tex
Some tuning for the TOPLAS camera-ready submission.
Add another phrase to acks.

@@
-\documentclass[format=acmsmall, review=true, screen=true]{acmart}
+\documentclass[format=acmsmall, screen=true]{acmart}
 
@@
 \author{Dominique~Devriese}
+\orcid{0000-0002-3862-6856}
 %\authornote{with author1 note}          %% \authornote is optional;
@@
 \ccsdesc[500]{Software and its engineering~Software verification}
+
+\keywords{capability machines, CHERI, local capabilities, well-bracketed control
+flow, stack encapsulation, logical relations, secure compilation}
 
@@
 Support for an STSM was received from COST Action EUTypes (CA15123).
-Dominique Devriese holds a Postdoctoral fellowship from the Research Foundation Flanders (FWO).
+Dominique Devriese held a Postdoctoral fellowship from the Research Foundation
+Flanders (FWO) during most of this research.
+This research was supported in part by the Research Foundation Flanders (FWO).
 \end{acks}
-
-%% Acknowledgments
-% \begin{acks}                            %% acks environment is optional
-%                                         %% contents suppressed with 'anonymous'
-%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%   %% acknowledge financial support and will be used by metadata
-%   %% extraction tools.
-%   This material is based upon work supported by the
-%   \grantsponsor{GS100000001}{National Science
-%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-%   conclusions or recommendations expressed in this material are those
-%   of the author and do not necessarily reflect the views of the
-%   National Science Foundation.
-% \end{acks}
 

### 2019-08-26 toplas/zip-sources.sh
[toplas] Add a small script for zipping the sources.

@@ -1,0 +1,2 @@
+rm toplas-sources.zip
+zip toplas-sources.zip acmart.cls ACM-Reference-Format.bst appendix.tex paper.tex preamble-illu.tex references.bib

### 2019-08-26 toplas/zip-sources.sh
[toplas] Add bibliography file, upload toplas-sources.zip

@@ -1,2 +1,2 @@
 rm toplas-sources.zip
-zip toplas-sources.zip acmart.cls ACM-Reference-Format.bst appendix.tex paper.tex preamble-illu.tex references.bib
+zip toplas-sources.zip acmart.cls ACM-Reference-Format.bst appendix.tex paper.tex preamble-illu.tex references.bib paper.bbl

### 2019-11-08 toplas/paper.tex
toplas: disable todonotes
added orcid for LB
Change journal name to TOPLAS
Address most TOPLAS publisher remarks.

@@
 % Metadata Information
-\acmJournal{TWEB}
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2010}
-\acmMonth{3}
-\copyrightyear{2009}
+\acmJournal{TOPLAS}
+\acmVolume{0}
+\acmNumber{0}
+\acmArticle{0}
+\acmYear{2019}
+\acmMonth{0}
+\copyrightyear{2019}
 %\acmArticleSeq{9}
@@
 
-\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+% \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+\usepackage[disable]{todonotes}
 % domi: IEEEtran.cls sets font size to 10bp, apparently.
@@
                                         %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\orcid{0000-0003-1320-0098}          %% \orcid is optional
 \affiliation{
@@
   We have made improvements to readability and completeness throughout the paper.
-  For instance, we have added an introduction to \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
+  For instance, we have added an introduction to \sectionname~\ref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.
   We highlight the following changes:
@@
 infinite address space and unbounded integers (see \sectionname~\ref{sec:discussion} for a discussion of these assumptions).
-
-We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
-We assume an infinite set of addresses $\Addrs$ and define machine words as either integers or capabilities of the form $((\perm,\gl),\var{base},\var{end},\addr)$.
-Such a capability represents the authority to execute permissions $\perm$ on the memory range $[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag $\gl$ indicating whether the capability is global or local.
-On a capability machine, there is no notion of pointers other than capabilities, so we will use the terms interchangeably.
-The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$), and read/write/execute ($\rwx$) permissions.
-Additionally, there are three special permissions: read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter ($\entry$), which we will explain below.
-\itoplas{The orderings of permissions and locality are displayed in \figurename~\ref{fig:perm-hier}.
-  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory which is why write-local is above write in the permission hierarchy.
-  We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
-
-\begin{wrapfigure}{R}{.4\textwidth}
-  \centering
-  \begin{tikzpicture}[main node/.style={}]
-    \node[main node] (7) {$\rwlx$};
-    \node[main node] (8) [below left of=7] {$\readwritel$};
-    \node[main node] (1) [below right of=7] {$\rwx$};
-    \node[main node] (2) [below right of=1] {$\exec$};
-    \node[main node] (3) [below right of=2] {$\entry$};
-    \node[main node] (4) [below left of=1] {$\readwrite$};
-    \node[main node] (5) [below right of=4] {$\readonly$};
-    \node[main node] (6) [below right of=5] {$\noperm$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (7) edge (8)
-    (7) edge (1)
-    (8) edge (4)
-    (1) edge (2)
-    (2) edge (3)
-    (2) edge (5)
-    (3) edge (6)
-    (1) edge (4)
-    (4) edge (5)
-    (5) edge (6);
-
-    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
-    \node[main node] (2l) [below of=1l] {$\local$};
-
-    \path[every node/.style={font=\sffamily\small}]
-    (1l) edge (2l);
-  \end{tikzpicture}
-  \caption{Permission and locality hierarchy.}
-  \label{fig:perm-hier}
-\end{wrapfigure}
-
-We assume a finite set of register names $\RegName$.
-We define register files $\reg$ and memories $\ms$ as functions mapping register names resp.\ addresses to words.
-The state of the entire machine is represented as a configuration that is either a running state $\Phi \in \ExecConfs$ containing a memory and a register file or a failed or halted state where the latter is paired with the final state of memory.
-
-% We also define memory segments $\ms$:
-% parts of memory represented as partial functions mapping addresses to words.
-
-%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
-
-The machine's instruction set is rather basic.
-Instructions $i$ include relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$), and move ($\move{}{}$, copies words between registers) instructions.
-Also familiar are load and store instructions for reading from and writing to memory ($\load{}{}$ and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and $\minus{}{}{}$, operating only on numbers).
-There are three instructions for modifying capabilities: $\lea{}{}$ (modifies the current address), $\restricttwo{}{}$ (modifies the permission and local/global tag), and $\subseg{}{}{}$ (modifies the range of a capability).
-Importantly, these instructions take care that the resulting capability always carries less authority than the original (e.g.\ $\restricttwo{}{}$ will only weaken a permission according to the hierarchy in \figurename~\ref{fig:perm-hier}).
-Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and $\geta{}{}$ provide access to a capability's permissions, local/global tag, base, end and current address, respectively.
 
@@
 \end{figure}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[main node/.style={}]
+    \node[main node] (7) {$\rwlx$};
+    \node[main node] (8) [below left of=7] {$\readwritel$};
+    \node[main node] (1) [below right of=7] {$\rwx$};
+    \node[main node] (2) [below right of=1] {$\exec$};
+    \node[main node] (3) [below right of=2] {$\entry$};
+    \node[main node] (4) [below left of=1] {$\readwrite$};
+    \node[main node] (5) [below right of=4] {$\readonly$};
+    \node[main node] (6) [below right of=5] {$\noperm$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (7) edge (8)
+    (7) edge (1)
+    (8) edge (4)
+    (1) edge (2)
+    (2) edge (3)
+    (2) edge (5)
+    (3) edge (6)
+    (1) edge (4)
+    (4) edge (5)
+    (5) edge (6);
+
+    \node[main node] (1l) [right of=1, xshift=2cm] {$\glob$};
+    \node[main node] (2l) [below of=1l] {$\local$};
+
+    \path[every node/.style={font=\sffamily\small}]
+    (1l) edge (2l);
+  \end{tikzpicture}
+  \caption{Permission and locality hierarchy.}
+  \label{fig:perm-hier}
+\end{figure}
+
+We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
+We assume an infinite set of addresses $\Addrs$ and define machine words as either integers or capabilities of the form $((\perm,\gl),\var{base},\var{end},\addr)$.
+Such a capability represents the authority to execute permissions $\perm$ on the memory range $[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag $\gl$ indicating whether the capability is global or local.
+On a capability machine, there is no notion of pointers other than capabilities, so we will use the terms interchangeably.
+The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$), and read/write/execute ($\rwx$) permissions.
+Additionally, there are three special permissions: read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter ($\entry$), which we will explain below.
+\itoplas{The orderings of permissions and locality are displayed in \figurename~\ref{fig:perm-hier}.
+  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory which is why write-local is above write in the permission hierarchy.
+  We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}
+
+We assume a finite set of register names $\RegName$.
+We define register files $\reg$ and memories $\ms$ as functions mapping register names resp.\ addresses to words.
+The state of the entire machine is represented as a configuration that is either a running state $\Phi \in \ExecConfs$ containing a memory and a register file or a failed or halted state where the latter is paired with the final state of memory.
+
+% We also define memory segments $\ms$:
+% parts of memory represented as partial functions mapping addresses to words.
+
+%\lau{we don't really make the distinction between memory and memory segments in the rest of the paper, so maybe we should leave this out?}
+
+The machine's instruction set is rather basic.
+Instructions $i$ include relatively standard jump ($\jmp{}$), conditional jump ($\jnz{}{}$), and move ($\move{}{}$, copies words between registers) instructions.
+Also familiar are load and store instructions for reading from and writing to memory ($\load{}{}$ and $\store{}{}$) and arithmetic operators ($\lt{}{}{}$ (less than), $\plus{}{}{}$ and $\minus{}{}{}$, operating only on numbers).
+There are three instructions for modifying capabilities: $\lea{}{}$ (modifies the current address), $\restricttwo{}{}$ (modifies the permission and local/global tag), and $\subseg{}{}{}$ (modifies the range of a capability).
+Importantly, these instructions take care that the resulting capability always carries less authority than the original (e.g.\ $\restricttwo{}{}$ will only weaken a permission according to the hierarchy in \figurename~\ref{fig:perm-hier}).
+Finally, the instruction $\isptr{}{}$ tests whether a word is a capability or a number and instructions $\getp{}{}$, $\getl{}{}$, $\getb{}{}$, $\gete{}{}$ and $\geta{}{}$ provide access to a capability's permissions, local/global tag, base, end and current address, respectively.
 
@@
     \end{itemize}
-    If we can show
+    If we can show\footnote{We ignore step indexes in this proof sketch.}
     \begin{equation}
@@
   \end{equation}
-  \footnote{We ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
+  then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
 

### 2019-11-15 presentations/aarhus-14-11-2019/main.tex
Add PhD defence slides

\documentclass[14pt]{beamer}
\setbeamertemplate{footline}[frame number]

\usepackage{appendixnumberbeamer}
\usepackage{xcolor}
\usepackage[official]{eurosym}
\input{preamble}
\input{illu-preamble}

\usepackage{pgfpages}
\setbeameroption{hide notes}
%\setbeameroption{show notes on second screen=right}
\usepackage{soul}

\usepackage[numbers,square]{natbib}
\usepackage{bibentry}
\nobibliography*
\bibliographystyle{plain}

\definecolor{myred}{RGB}{192,58,45}
\definecolor{mypurple}{RGB}{180,52,202}
\definecolor{myblue}{RGB}{51,51,179}
\definecolor{mylblue}{RGB}{52,178,204}
\definecolor{mygreen}{RGB}{66,192,103}
\usepackage{tikzsymbols}

\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{calc}
\usetikzlibrary{arrows,shapes}

\tikzset{
    invisible/.style={opacity=0,text opacity=0},
    visible on/.style={alt=#1{}{invisible}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} 
    },
    beameralert/.style={alt=<#1>{fill=red!30,rounded corners}{},anchor=base},
    BeamerAlert/.style={alt=#1{fill=red!30,rounded corners}{},anchor=base}
}
\tikzset{capa/.style={decoration={markings,mark=at position 0 with %
      {\fill[black] circle(1.5pt);}}}}

\def\rotatechar#1{\reflectbox{#1}}
\newcommand{\capchar}{\euro{}}

\usepackage{graphicx}
\usepackage{dashrule}

\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape=true,
  literate={\$}{\$}{1},
  keywordstyle=\color{mylblue},
  tabsize=2
}

\lstdefinelanguage
   [x64]{Assembler}     % add a "x64" dialect of Assembler
   [x86masm]{Assembler} % based on the "x86masm" dialect
   % with these extra keywords:
   {morekeywords={CDQE,CQO,CMPSQ,CMPXCHG16B,JRCXZ,LODSQ,MOVSXD, %
                  POPFQ,PUSHFQ,SCASQ,STOSQ,IRETQ,RDTSCP,SWAPGS, %
                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp, %
                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b, %
                  r10,r10d,r10w,r10b,r11,r11d,r11w,r11b, %
                  r12,r12d,r12w,r12b,r13,r13d,r13w,r13b, %
                  r14,r14d,r14w,r14b,r15,r15d,r15w,r15b}} % etc.

\tikzset{
  lock/.pic = {
    \draw[even odd rule,fill=gray] (.1,.2) circle[radius=.08]
    (.1,.2) circle[radius=.06];
    \draw[fill=gray] (0,0) rectangle (.2,.2);

    \draw (.086,.054) rectangle (.114,.14);
    \draw (.1,.13) circle[radius=.03];

    \draw[fill=white,draw=none](.096,.064) rectangle (.104,.13);
    \draw[fill=white,draw=none] (.1,.13) circle[radius=.02];
  }
}

\tikzset{capa/.style={decoration={markings,mark=at position 0 with %
      {\fill[black] circle(1.5pt);}}}}

\tikzset{capar/.style={decoration={markings,mark=at position 0 with %
      {\fill[myred] circle(1.5pt);}}}}

% #1 number of teeths
% #2 radius intern
% #3 radius extern
% #4 angle from start to end of the first arc
% #5 angle to decale the second arc from the first
% #6 inner radius to cut off

\newcommand{\gear}[6]{%
  (0:#2)
  \foreach \i [evaluate=\i as \n using {\i-1)*360/#1}] in {1,...,#1}{%
    arc (\n:\n+#4:#2) {[rounded corners=1.5pt] -- (\n+#4+#5:#3)
    arc (\n+#4+#5:\n+360/#1-#5:#3)} --  (\n+360/#1:#2)
  }%
  (0,0) circle[radius=#6] 
}

\usepackage{tikz-cd}
\pgfdeclarelayer{bg}
\pgfdeclarelayer{mg}
\pgfdeclarelayer{fg}
\pgfdeclarelayer{ffg}
\pgfsetlayers{bg,mg,main,fg,ffg}

% \draw <coord> pic {lock};

% \AtBeginSection[]
% {
%   \begin{frame}<beamer>
%     \frametitle{Overview}
%     \tableofcontents[currentsection]
%   \end{frame}
% }
\tikzset{
  capcoin/.pic={
        code={\tikzset{scale=3/10}
          \node[circle,draw,inner sep=1pt,shading=axis,left color=yellow,right color=yellow!10!white,shading angle=-135] (coin) {\hspace{-1pt}\phantom\capchar{}};
        }}
      }

\tikzset{
  smallcapcoin/.pic={
        code={\tikzset{scale=3/10}
          \node[circle,draw,inner sep=1pt,shading=axis,left color=yellow,right color=yellow!10!white,shading angle=-135] (coin) {\tiny\hspace{-1pt}\capchar{}};
        }}
      }

\tikzset{
  smallcapcoing/.pic={
        code={\tikzset{scale=3/10}
          \node[circle,draw,inner sep=1pt,shading=axis,left color=green,right color=green!10!white,shading angle=-135] (coin) {\tiny\hspace{-1pt}\capchar{}};
        }}
      }

\tikzset{
  smallcapcoinb/.pic={
        code={\tikzset{scale=3/10}
          \node[circle,draw,inner sep=1pt,shading=axis,left color=blue,right color=blue!10!white,shading angle=-135] (coin) {\tiny\hspace{-1pt}\capchar{}};
        }}
      }

%\input{../illustrations/preamble}
%\input{loc_preamble}
      \title{Formal Reasoning\\\large{about}\\\Large{Capability Machines}}
%\subtitle{PhD Defence}
\author{Lau Skorstengaard}
\date{PhD Defence\\November 14, 2019}
\titlegraphic{\includegraphics[width=5cm]{au_logo}}

\begin{document}
\beamertemplatenavigationsymbolsempty
\bibliography{ref}
\maketitle
% Tell plan for talk, i.e.
% In this presentation, I am going to tell you about my thesis: Formal Reasoning about capability machines.
% The presentation is split in two parts.
% In the first part, I will motivate why we need to care about this work, and I will set the scene and give you a basic understanding about some concepts you will need to know about.
% In the second part, I will present the contents of the thesis it self and the papers that constitute the thesis.
% And I am going to walk you through the results of those papers.

\begin{frame}[label=motivation]
  \frametitle{Programming}
  \centering

  \begin{tikzpicture}
    \begin{scope}[visible on=<+->]
      \node[rotate=30,align=center,rectangle, rounded corners,draw] at (-4,2) {High-level};
      \draw (-5.5,0) edge[dashed] (5.5,0);
      \node[rotate=30,align=center,rectangle,rounded corners,draw] at (-4,-2) {Low-level};
    \end{scope}

    \begin{scope}[visible on=<+->]
      \node[alice,monitor,mirrored,minimum width=2cm] at (4,2) {};
      \draw[visible on=<.-.(4)>]node[bob,monitor,mirrored,minimum width=2cm] at (4,-2) {};
    \end{scope}

    \begin{scope}[visible on=<+->]
      \draw[thick] (0,.5) -- (-1,.5) -- (-1,3) -- (.5,3) -- (1,2.5) -- (1,2);
      \draw[rounded corners, thick] (.5,3) -- (.5,2.5) -- (1,2.5);

      \draw[dashed,draw=black!25!white] (0,.5) rectangle (1,2);
      \pattern[pattern=north east lines,pattern color=black!25!white] (0,.5) rectangle (1,2);

      \node[anchor=north west, align=left, font=\tiny\ttfamily] at (-1,3) {main:\\\ \ x := 5\\\ \ ...};

      \coordinate (b1) at (-1,.5);
      \coordinate (b2) at (1,3);
      \node[fit=(b1) (b2),rectangle, inner sep=0pt] (a-file-1) {};
    \end{scope}

    \begin{scope}[visible on=<+->]
      \draw[fill=white,draw=white] (.5,-3.5) rectangle (1.5,-2) node[pos=.5] (b-prog) {};
      \draw[dashed,draw=black!50!white] (.5,-3.5) -- (.5,-2) -- (1.5,-2);
      \draw[thick] (.5,-3.5) -- (1.5,-3.5) -- (1.5,-2);
      \node[anchor=north west, align=left, font=\tiny\ttfamily] at (.5,-2) {1\\10\\001\\1010\\11101};
    \end{scope}

    \begin{scope}[visible on=<+->]
      \node[anchor=north west, align=left, font=\tiny\ttfamily] at (-1,-.5) {0\\01\\101\\0010\\110010\\001001\\110010\\001010\\111001};
      \draw[thick] (-1,-.5) -- (-1,-3) -- (0,-3);
      \draw[dashed,draw=black!25!white] (0,-3) -- (0,-1.5) -- (1,-1.5);
      \draw[thick](1,-1.5) -- (1,-1) -- (.5,-.5) -- (-1,-.5);
      \draw[rounded corners, thick] (1,-1) -- (.5,-1) -- (.5,-.5);

      \begin{pgfonlayer}{bg}
        \begin{scope}[visible on=<.->]
        \draw[dashed,draw=black!25!white] (0,-3) rectangle (1,-1.5);
        \pattern[pattern=north east lines,pattern color=black!25!white] (0,-3) rectangle (1,-1.5);
      \end{scope}
      \end{pgfonlayer}

      \coordinate (a1) at (-1,-.5);
      \coordinate (a2) at (1,-3);
      \node[fit=(a1) (a2),rectangle, inner sep=0pt] (a-file-2) {};
      \draw[visible on=<.>] (a-file-1.west) edge[-latex,bend right] ([yshift=-1em]a-file-2.north west);
    \end{scope}

    \begin{scope}[visible on=<+->]
      \draw[shorten >=3pt, shorten <=3pt,-latex] (.5,-2) -- (0,-1.5);
    \end{scope}

    \begin{scope}[visible on=<+->]
      \node[bob,evil,monitor,mirrored,minimum width=2cm] at (4,-2) {};
      \node[devil,mirrored] at (b-prog) {};
    \end{scope}

  \end{tikzpicture}
\end{frame}

% \begin{frame}[label=title-hl]
%   % consider changing HL style.
%   \begin{center}
%     \Huge\usebeamercolor[fg]{frametitle}{
%     \tikz[remember picture] \node[inner sep=0pt] (frl) {};Formal Reasoning\tikz[remember picture] \node[inner sep=0pt] (frr) {};\\
%     about\\
%     \tikz[remember picture] \node[inner sep=0pt] (c) {};Capability\tikz[remember picture] \node[inner sep=0pt] (mid) {}; \tikz[remember picture] \node[inner sep=0pt] (mid2) {};Machines\tikz[remember picture] \node[inner sep=0pt] (m) {};
%     }
%   \end{center}
%   % Highlight each part of the title.
%   \begin{tikzpicture}[remember picture, overlay]
%     \coordinate (frll) at ([xshift=0.2cm,yshift=0.6cm]frl);
%     \coordinate (frrr) at ([xshift=-0.2cm,yshift=0.1cm]frr);
%     \draw<2> node[fit=(frll) (frrr), thick, rectangle, draw=red, rounded corners, inner sep=17pt] {};
%     \coordinate (cc) at ([xshift=0.3cm,yshift=0.6cm]c);
%     \coordinate (mm) at ([xshift=-0.2cm,yshift=0.6cm]m);
%     \coordinate (midd) at ([xshift=-0.3cm,yshift=0.1cm]mid);
%     \coordinate (midd2) at ([xshift=0.3cm,yshift=0.1cm]mid2);
%     \draw<3>  node[fit=(cc) (mm) (midd), thick, rectangle, draw=red, rounded corners, inner sep=17pt] {};
%     \draw<4>  node[fit=(mm) (midd2), thick, rectangle, draw=red, rounded corners, inner sep=17pt] {};
%     \draw<5>  node[fit=(cc) (midd), thick, rectangle, draw=red, rounded corners, inner sep=17pt] {};
%   \end{tikzpicture}
% \end{frame}

% Highlight Machine

% \againframe<4>{title-hl}

\begin{frame}<3-18>[label={cap-machine-fig}]
  \frametitle<1>{What is a machine?}
  \frametitle<2-18>{What is a computer?} 
  \frametitle<19->{What is a capability machine?}
  % Big picture of a computer?

  % Computation unit (me)
  % I can perform simple computations like adding numbers together.
  %% Thought bubble with numbers.
  %% Adding them together.
  % I can only keep so many numbers in my mind, so I have a notepad where I can write them and read them back when I need them.
  % What computations do I perform? I read them from the notepad - this is what we know as a program.
  % There may be many different authors of these programs. 
  % What prevents programs from interfering with each other maliciously?
  % Introduce capabilities...
  \begin{center}
  \begin{tikzpicture}
    \def\figspacing{2cm}
    % CPU-MAN

    \draw[visible on=<2-25>] node[businessman,monogramtext=CPU,mirrored,minimum size=1.5cm] (cpu) {};

    % REGISTERS
    \begin{scope}[visible on=<2-25>]
    \coordinate[right=of cpu,yshift=3.5em, xshift=-4pt] (reg0);
    \draw[visible on=<4-9>] node (reg0w1) at (reg0) {\alert<8-9>{5}};
    \draw[visible on=<4-9>] (cpu.40) -- (reg0w1);
    \draw[visible on=<19->] (reg0) pic {capcoin};
    \draw[visible on=<19->] (cpu.40) -- (reg0w1);

    \coordinate[right=of cpu,yshift=2em] (reg1);
    \draw[visible on=<4-9>] node (reg1w1) at (reg1) {1};
    \draw[visible on=<4->] (cpu.30) -- (reg1w1);
    \draw[visible on=<10-20>] node (reg1w5) at (reg1) {\alert<11>{5}};
    \draw[visible on=<21->] node (reg1w5) at (reg1) {\alert<20-23>{5}};

    \coordinate[right=of cpu,yshift=0.5em, xshift=4pt] (reg2);
    \draw[visible on=<4-9>] node (reg2w2) at (reg2) {2};
    \draw[visible on=<4-9>] (cpu.20) -- (reg2w2);
    \draw[visible on=<12->] node (reg2w3) at (reg2) {3};
    \draw[visible on=<12->] (cpu.20) -- (reg2w3);

    \coordinate[right=of cpu, yshift=-1em,xshift=8pt] (reg3);
    \draw[visible on=<6-9>] node (reg3w3) at (reg3) {\alert<8-9>{3}};
    \draw[visible on=<6-9>] (cpu.10) -- (reg3w3);
    \draw[visible on=<14->] node (reg3w8) at (reg3) {8};
    \draw[visible on=<14->] (cpu.10) -- (reg3w8);
  \end{scope}
    % OPERATION
    \begin{scope}[visible on=<5-6>]
    \node (plus) at ($(reg1)!0.5!(reg2)$) {\scriptsize +};
    \node (assign) at ($(reg2)!0.5!(reg3)$) {\scriptsize =};
    \end{scope}
    \begin{scope}[visible on=<13-14>]
    \node (plus) at ($(reg1)!0.5!(reg2)$) {\scriptsize +};
    \node (assign) at ($(reg2)!0.5!(reg3)$) {\scriptsize =};
    \end{scope}

    % CAPABILITY DYNAMIC CHECKER
    \draw[visible on=<20-26>] node[maninblack, minimum size=1.5cm,left=\figspacing*1.5 of cpu] (check1) {};

    % MEMORY
    \def\notepadsize{1.75cm}
    \def\papersize{1.45cm}
    \def\paperpadding{0.1cm}
    \def\numberoflines{9}

    \begin{scope}[visible on=<7-26>]
    \node[left=\figspacing*.5 of check1,rounded corners, minimum width=\notepadsize, minimum height=\notepadsize*1.414,draw,shading = axis, left color=brown, right color=brown!80!white,shading angle=-135] (mem) {};
    \node[minimum width=\papersize, minimum height=\papersize*1.414,draw,fill=white] (paper) at (mem) {};
    \begin{scope}[shift=(mem)]
    \foreach \ln in {1, ..., \numberoflines}{   
      \draw (\paperpadding-\papersize/2,{\ln*(\papersize*1.414-\paperpadding*2)/\numberoflines-\papersize*1.414/2}) -- (\papersize/2-\paperpadding, {\ln*(\papersize*1.414-\paperpadding*2)/\numberoflines-\papersize*1.414/2});
    }

    \end{scope}
    \coordinate (clip) at ($(mem) - (\papersize/4,-\papersize*1.414/2+\paperpadding/2)$);
    \begin{scope}[shift=(clip)]
    \filldraw[fill=gray,draw=gray] (0,0) -- (\papersize/16,0) -- (\papersize/16,\papersize/16) arc (90:180:\papersize/16);
    \filldraw[fill=gray,draw=gray] (\papersize/16,0) rectangle (7*\papersize/16,\papersize/16);  
    \filldraw[fill=gray,draw=gray] (7*\papersize/16,0) -- (8*\papersize/16,0) arc (0:90:\papersize/16) -- cycle;
    \filldraw[fill=gray,draw=gray] (3*\papersize/16,\papersize/16) -- (5*\papersize/16,\papersize/16) arc (0:180:\papersize/16);
    \draw (0,0) -- (\papersize/2,0) edge[out=90,in=0] (\papersize/2-\papersize/16,\papersize/16);
    \draw (\papersize/2-\papersize/16,\papersize/16) edge (\papersize/2-3*\papersize/16,\papersize/16);
    \draw (\papersize/2-3*\papersize/16,\papersize/16) arc (0:180:\papersize/16);
    \draw (\papersize/2-5*\papersize/16,\papersize/16) edge (\papersize/2-7*\papersize/16,\papersize/16);
    \draw (\papersize/2-7*\papersize/16,\papersize/16) edge[out=180,in=90] (0,0);
    \end{scope}
    \end{scope}

    % STORE NUMBER
    \coordinate (memright) at ($(mem) + (\notepadsize/2 + 5,0)$);
    \draw[visible on=<8-9>] (cpu.west) edge[-stealth] node[above] {Write \alert{3} on line \alert{5}} (memright);

    \coordinate (line4) at ($(mem)+(0,{.5*(\papersize*1.414-\paperpadding*2)/\numberoflines})$);
    \draw[visible on=<9-26>] node at (line4) {\tiny 3};

    % LOAD NUMBER
    \draw[visible on=<11>] ([yshift=.5cm]cpu.west) edge[-stealth] node[above] {Read from line \alert{5}} ([yshift=.5cm]memright);
    \draw[visible on=<11>] ([yshift=-.5cm]memright) edge[-stealth] node[above] {3} ([yshift=-.5cm]cpu.west);

    % READ PROGRAM
    \draw[visible on=<15>] ([yshift=.5cm]cpu.west) edge[-stealth] node[above] {What's next?} ([yshift=.5cm]memright);
    \draw[visible on=<15>] ([yshift=-.5cm]memright) edge[-stealth] node[above] {Add some numbers!} ([yshift=-.5cm]cpu.west);

    % PROGRAM AUTHORS
    \def\authordist{1cm}
    \coordinate (abovemem) at ($(mem)+(0,\notepadsize*1.414/2+2)$);

    \begin{scope}[visible on=<16-26>]
    \draw node[builder, monitor, minimum size=.75cm] (authormid) at ([xshift=.5cm]$(mem) + (0,\notepadsize*1.414/2+\authordist)$) {};
    \draw ([yshift=-2pt]authormid.south) edge[-stealth] (abovemem);
    \end{scope}

    \begin{scope}[visible on=<17-26>]
    \draw[visible on=<17>] node[priest, monitor,minimum size=.75cm,left=\authordist/1.5 of authormid] (authorleft) {};    
    \draw ([yshift=-2pt]authorleft.south) edge[-stealth] ($(abovemem) + (-\notepadsize/2+\notepadsize/16,0)$);

    \draw node[cowboy, monitor,mirrored, minimum size=.75cm,right=\authordist/1.5+.5cm of authormid] (authorright) {};
    \draw ([yshift=-2pt]authorright.south west) edge[-stealth] ($(abovemem) + (\notepadsize/2-\notepadsize/16,0)$);
    \end{scope}

    \draw[visible on=<18-26>] node[priest, evil, monitor,minimum size=.75cm,left=\authordist/1.5 of authormid] (authorleftevil) {};

    % INTRODUCING CAPABILITIES
    % TODO consider changing load arrow to text bobble
    \draw[visible on=<21-24>] (cpu.west) edge[-stealth] node[above] (label){\small{Read from line \alert{5}}} (check1.east); 
    \node[visible on=<22>, ellipse callout, draw,yshift=.3cm,callout absolute pointer={(check1.mouth)}] at ([xshift=.5cm]check1.north east) {No!};
    \draw[visible on=<23-24>] ([yshift=-.75cm]label) pic {capcoin};
    \node[visible on=<24>, ellipse callout, draw,yshift=.3cm,callout absolute pointer={(check1.mouth)}] at ([xshift=.5cm]check1.north east) {3!};

    % COMPARTMENTALIZATION
    % \begin{scope}[visible on=<24>]
    %   \draw ([yshift=6pt]authorleftevil.north west) pic (c1) {smallcapcoin};
    %   \draw ([yshift=14pt,xshift=6pt]authorleftevil.north west) pic (c2) {smallcapcoin};
    %   \draw ([yshift=18pt,xshift=-4pt]authorleftevil.north west) pic (c3){smallcapcoin};

    %   \draw ([yshift=6pt]authormid.north west) pic (c4) {smallcapcoing};
    %   \draw ([yshift=14pt,xshift=6pt]authormid.north west) pic (c5) {smallcapcoing};
    %   \draw ([yshift=18pt,xshift=-4pt]authormid.north west) pic (c6) {smallcapcoing};

    %   \draw ([yshift=6pt]authorright.north east) pic (c7) {smallcapcoinb};
    %   \draw ([yshift=14pt,xshift=-6pt]authorright.north east) pic (c8) {smallcapcoinb};
    %   \draw ([yshift=18pt,xshift=4pt]authorright.north east) pic (c9){smallcapcoinb};
    % \end{scope}

    \begin{scope}[visible on=<26->]
      \draw (0.5,2.5) node {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (r1) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_1$};
      \coordinate (r2) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_2$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};

      \draw[visible on=<26-27>] (r1) pic {capcoin};
      \node at (r2) {5};
      \node at (r3) {3};
      \node at (r4) {8};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[visible on=<27->,shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
      \end{scope}
      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
      \foreach \x in {.25,.5,...,5}
      {
        \draw (0,\x) rectangle (2.5,\x+.25);
      };
      \node at (1.25,6) {Memory};
    \end{scope}

    \begin{scope}[visible on=<28->, shift=(shifter2)]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,4.25) node (e) [black,midway] {};
      \draw[visible on=<29->,decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.6,1) -- (2.6,4.25) node [black,midway] {};
      \draw[visible on=<28-29>,capa] (r1) edge[-latex,postaction={decorate},out=180,in=0] (e);
      \draw[visible on=<30>,capa] (r1) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rwx}} (e);
    \end{scope}
    % TODO Do I want to demonstrate some instructions?

    % Give programmers different capabilities (colors?)
  \end{tikzpicture}
  \end{center}
\end{frame}

% Highlight capability
%\againframe<5>{title-hl}
\begin{frame}
  \frametitle{What is a capability?}
\onslide<2>{  Generally, a capability is 
  \begin{center}
    \huge
    \textit{an unforgeable\\token of authority}
  \end{center}
}  \note{Here, we prohibit arbitrary access to the notepad and require that access only can happen if you have the correct capability.}
\end{frame}

%\againframe<3>{title-hl}
\againframe<19->{cap-machine-fig}
% Highlight capability machine

\newsavebox{\assem}
\begin{lrbox}{\assem}
\begin{minipage}{.3\textwidth}
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
main:
	.cfi_startproc
# BB#0:
	pushq	%rbp
.Ltmp0:
	.cfi\_def\_cfa\_offset 16
.Ltmp1:
	.cfi\_offset %rbp, -16
	movq	%rsp, %rbp
\end{lstlisting}
\end{minipage}
\begin{minipage}{.3\textwidth}
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
.Ltmp2:
	.cfi\_def\_cfa\_register %rbp
	subq	$\mbox{\textdollar}$16, %rsp
	movabsq	$\mbox{\textdollar}$.L.str, %rdi
	movl	$\mbox{\textdollar}$0, -4(%rbp)
	movb	$\mbox{\textdollar}$0, %al
	callq	printf
	xorl	%ecx, %ecx
	movl	%eax, -8(%rbp)
\end{lstlisting}
\end{minipage}
\begin{minipage}{.3\textwidth}
\vspace{1em}
\begin{lstlisting}[basicstyle=\tiny\ttfamily]
	movl	%ecx, %eax
	addq	$\mbox{\textdollar}$16, %rsp
	popq	%rbp
	retq
.Lfunc\_end0:
	.size	main, .Lfunc\_end0-main
	.cfi\_endproc
\end{lstlisting}
\end{minipage}

\end{lrbox}

\newsavebox{\hello}
\begin{lrbox}{\hello}
\begin{lstlisting}[basicstyle=\tiny\ttfamily, language=C, keywordstyle=\bfseries]
#include <stdio.h>
int main()
{
  int t = 5;
  printf("Hello, World!");
  return 0;
}
\end{lstlisting}
\end{lrbox}

\againframe<2>{motivation}
% \begin{frame}
%   \frametitle{Layers of abstraction}
%   \note{Why do we want capabilities.}
%   \note{Enforce abstractions.}
%   \begin{center}
%   \begin{tikzpicture}

%     \node[visible on=<1->,inner sep=0pt,anchor=north] (hardware) at (0,0) {\includegraphics[width=.8\textwidth]{hardware2}};  
%     \draw[visible on=<2->,dashed] (-5.5,.2) -- (5.5,.2);
%     % (TODO) add label to each layer (hardware, assembly, high-level language).
% %    \node[rotate=45,rectangle, draw=black,shading=axis,left color=blue,right color=blue!10!white,shading angle=-135,rounded corners, inner sep=3pt] (hw-label) at (hardware.west) {\scriptsize Hardware};

%     \node[visible on=<2->,inner sep=0pt,anchor=south] (mid) at ($(hardware.north) + (0,.4)$) {\usebox{\assem}};  
% %    \node[rotate=45,rectangle, draw=black,shading=axis,left color=blue,right color=blue!10!white,shading angle=-135,rounded corners, inner sep=3pt] (hw-label) at (mid.west) {\scriptsize x86 assembly};
%     \draw[visible on=<3->,dashed] ($(mid.north) + (-5.5,.2)$) -- ($(mid.north) + (5.5,.2)$);

%     \node[visible on=<3->,inner sep=0pt,anchor=south] (top) at ($(mid.north) + (0,.4)$) {\usebox{\hello}};  
% %    \node[rotate=45,rectangle, draw=black,shading=axis,left color=blue,right color=blue!10!white,shading angle=-135,rounded corners, inner sep=3pt] (hw-label) at (top.west) {\scriptsize C};

%     \draw[visible on=<4>] ($(top)-(2,0)$) edge[-stealth,bend right,very thick,align=center] node[left] {Translation} ($(mid.north)-(2,0)$);

%     \begin{pgfonlayer}{ffg}
%       \node (low-level) at ($(hardware) + (5.5,-.5)$) {Low};
%       \node (high-level) at ($(top) + (5.5,.5)$) {High};
%       \draw (low-level) edge[very thick] node[above,rotate=-90] (scale-label) {\scriptsize level of abstraction} (high-level);
%     \end{pgfonlayer}
%     \begin{pgfonlayer}{fg}
%       \draw node[rectangle, fit=(low-level)(high-level)(scale-label),draw=black,shading=axis,left color=blue!50!white,right color=blue!10!white,shading angle=-135,rounded corners, inner sep=1pt] {};
%     \end{pgfonlayer}
%   \end{tikzpicture}
% \end{center}
% \end{frame}

\begin{frame}[fragile]
  \frametitle<1-12>{Program execution}
  \frametitle<13>{Local-state encapsulation}
  \frametitle<14->{Not well-bracketed control-flow}
  \note{If we can make this happen on lower level, then we have an abstraction failure.}
  \begin{columns}
    \begin{column}{0.5\textwidth}
\begin{lstlisting}[language=C, escapechar=!,basicstyle=\footnotesize\ttfamily]
 void a()
 {
   !\tikz[remember picture] \node[] (pc1) {};!int x = 5;
   !\tikz[remember picture] \node[] (pc2) {};!b();
   !\tikz[remember picture] \node[] (pc3) {};!...
   !\tikz[remember picture] \node[] (pc4) {};!b();
   !\tikz[remember picture] \node[] (pc5) {};!return;!\tikz[remember picture] \node[] (pc5r) {};!
 }

!\tikz[remember picture] \node[] (a) {};!void b()       !\tikz[remember picture] \node[] (bt) {};! 
 {
   !\tikz[remember picture] \node[] (pc6) {};!...
   !\tikz[remember picture] \node[] (pc7) {};!return;!\tikz[remember picture] \node[] (ret) {};!      
 }              !\tikz[remember picture] \node[] (bb) {};!

\end{lstlisting}
% Mention that our work support high-order
% No good definition. We appeal to you intuitive definition
\newcommand{\myarrow}[2]{\draw<#1> ($#2 + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} #2;}
\newcommand{\myarrowred}[2]{\draw<#1> ($#2 + (-0.5,0)$) edge[-stealth,thick,draw=red] node {} #2;}
      \begin{tikzpicture}[remember picture, overlay]
%        \draw<1> node { };
        \myarrow{3}{(pc1)}
        \myarrow{4}{(pc2)}
        \myarrow{5}{(a)}
        \myarrow{6}{(pc7)}
        \myarrow{7}{(pc3)}
        \myarrow{8}{(pc4)}
        \myarrow{9}{(a)}
        \myarrow{10}{(pc7)}
        \myarrow{11}{(pc5)}

         \draw<13>[decorate,decoration={brace,amplitude=6pt,raise=4pt},yshift=0pt] ($(bt) + (0,0.2)$) -- ($(bb) + (0,-0.2)$) node[xshift=0.2cm] (brace) [midway] {};
         \draw<13> node[right = -0cm of brace, text width = 3cm] {Function \texttt{b} cannot access variable \texttt{x}};

        \myarrow{14}{(pc4)}
        \myarrow{15}{(a)}
        \myarrow{16}{(pc7)}

        \myarrowred{17}{(pc3)}
        \myarrowred{18}{(pc3)}
        \node[visible on=<18>,below right=.75cm of pc5r,align=center] (wbcferr) {Should have\\returned here!};
        \draw[visible on=<18->] (wbcferr.north) edge[-stealth,bend right] (pc5r);

      %   \draw<3> ($(pc4) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc4);
      %   \draw<4> ($(a) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (a);
      %   \draw<5> ($(pc6) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc6);
      %   \draw<6> ($(pc7) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc7);
      %   \draw<7-> node[below right = of ret] (w) {\huge ?};
      %   \draw<7-> (ret) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202},bend left] (w);
      %   \draw<8> ($(pc5) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc5);
      %   \draw<9> ($(pc3) + (-0.5,0)$) edge[-stealth,thick,draw={rgb:red,180;green,52;blue,202}] node {} (pc3);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{tikzpicture}
        \begin{scope}[visible on=<2->]
        \draw node (stack label) at (1.25,6) {Call stack};
        \stdstackstartclean[5.5]
        \draw (3,0) edge[-latex] (3,2);
        \begin{scope}
          \clip (-0.1,0) rectangle (2.6,1.35);
          \draw[fill=gray!50] (0,-0.2) rectangle (2.5,1.25) node[pos=.5] {\scriptsize{Lower stack frames}};
        \end{scope}
      \end{scope}
        \coordinate (shifter) at (0,1);
        \begin{scope}[shift=(shifter)]
        \draw[visible on=<2-11>, fill=white] (0,.25) rectangle (2.5,1.5) node[pos=.5] {\texttt{a()}};
        \draw[visible on=<3-11>] node (xvar) at (2,.75) {\texttt{\scriptsize{x}}};
        \draw[visible on=<5-6>,fill=white] (0,1.5) rectangle (2.5,2.75) node[pos=.5] {\texttt{b()}};
        \draw[visible on=<9-10>,fill=white] (0,1.5) rectangle (2.5,2.75) node[pos=.5] {\texttt{b()}};

        \draw[visible on=<13->, fill=white] (0,.25) rectangle (2.5,1.5) node[pos=.5] {\texttt{a()}};
        \draw[visible on=<13>] node (xvar) at (2,.75) {\texttt{\scriptsize{x}}};
        \draw[visible on=<13>,fill=white] (0,1.5) rectangle (2.5,2.75) node[pos=.5] (bframe) {\texttt{b()}};
        \draw[visible on=<13>] (bframe) edge[-stealth,thick,draw=red,bend left] node (midarrow) {} (xvar);
        \draw[visible on=<13>] ($(midarrow) + (-.15,.15)$) edge[thick,draw=red] ($(midarrow) - (-.15,.15)$);
        \draw[visible on=<13>] ($(midarrow) + (.15,.15)$) edge[thick,draw=red] ($(midarrow) - (.15,.15)$);

        \draw[visible on=<15-16>,fill=white] (0,1.5) rectangle (2.5,2.75) node[pos=.5] {\texttt{b()}};

      \end{scope}
      \end{tikzpicture}
    \end{column}
  \end{columns}
\end{frame}
\againframe<2>{motivation}
\begin{frame}
  \frametitle{A na\"ive capability calling convention}
    \centering
  \begin{tikzpicture}
    % recurrent parts
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (rstk) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_\stk$};
      \coordinate (r2) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_2$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \draw[fill=gray!50] (0,-.2) rectangle (2.5,1) node[pos=.5] {\scriptsize Lower stack frames};
      \end{scope} 
      \begin{scope}
        \clip (-.1,-1.4) rectangle (2.6,-.6);
        \draw[white] (0,-1.5) rectangle (2.5,-.5) node[pos=.5] (return) {};
      \end{scope}
      \node at (1.25,6) {Call stack};
      \draw[visible on=<1->,fill=white] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};
      \draw[visible on=<4>,fill=white] (0,2.5) rectangle (2.5,4) node[pos=.5] {\texttt{b()}};
      \draw[visible on=<4-5>,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      % B'S STACK POINTER
      \begin{scope}[visible on=<1>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwx}} (e);
      \end{scope}
      \begin{scope}[visible on=<6>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwx}} (e);
      \end{scope}

      % A'S STACK POINTER
      \begin{scope}[visible on=<2-4>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.9,2.5) -- (2.9,6) node[black,pos=0.4] (d) {};
        \begin{scope}
          \clip (2.5,2.4) rectangle (3.5,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.4,mirror},yshift=0pt] (2.9,2.5) -- (2.9,6) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rwx}} (d);
      \end{scope}

      % B'S MOVED STACK POINTER
      \begin{scope}[visible on=<2>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) {};
        \end{scope}
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwx}} (e);
      \end{scope}

      % B'S SEALED STACK POINTER
      \begin{scope}[visible on=<3-5>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) {};
        \end{scope}
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[below] (blabel) {} (e);
        \draw (blabel.south) pic {lock};
      \end{scope}

    \end{scope} %shift=(shifter2)
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Breaking local-state encapsulation}
    \centering
  \begin{tikzpicture}
    % recurrent parts
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (rstk) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_\stk$};
      \coordinate (r2) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_2$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \draw[fill=gray!50] (0,-.2) rectangle (2.5,1) node[pos=.5] {\scriptsize Lower stack frames};
      \end{scope}
      \begin{scope}
        \clip (-.1,-1.4) rectangle (2.6,-.6);
        \draw[white] (0,-1.5) rectangle (2.5,-.5) node[pos=.5] (return) {};
      \end{scope}

      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
      \node at (1.25,6) {Call stack};
      \draw[visible on=<1->,fill=white] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};
      \draw[visible on=<1-2>,fill=white] (0,2.5) rectangle (2.5,4) node[pos=.5] (b-program) {\texttt{b()}};
      \draw[visible on=<1-2>] node[devil,minimum size=.5cm,anchor=west] at (b-program.east) {};a
      \draw[visible on=<1-2>] node[devil,mirrored,minimum size=.5cm,anchor=east] at (b-program.west) {};
      \draw[visible on=<1-2>,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      % A'S STACK POINTER
      \begin{scope}[visible on=<1>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.9,2.5) -- (2.9,6) node[black,pos=0.4] (d) {};
        \begin{scope}
          \clip (2.5,2.4) rectangle (3.5,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.4,mirror},yshift=0pt] (2.9,2.5) -- (2.9,6) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rwx}} (d);
      \end{scope}

      % B'S SEALED STACK POINTER
      \begin{scope}[visible on=<1-2>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) {};
        \end{scope}
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[below] (blabel) {} (e);
        \draw (blabel.south) pic {lock};
      \end{scope}

      % A'S STORED STACK POINTER
      \begin{scope}[visible on=<2-5>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (4.2,2.5) -- (4.2,6) node[black,pos=0.4] (d) {};
        \begin{scope}
          \clip (3.5,2.4) rectangle (4.5,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.4,mirror},yshift=0pt] (4.2,2.5) -- (4.2,6) {};
        \end{scope}
        \draw[capa] ([yshift=.5cm]reg-file.north west) edge[-latex,postaction={decorate},out=180,in=0] node[above left] {\scriptsize\textsc{rwx}} (d);
      \end{scope}

      % B'S STACK POINTER
      \begin{scope}[visible on=<3-4>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwx}} (e);
      \end{scope}

      % STACK FRAMES
      \draw[visible on=<4>,fill=white] (0,1) rectangle (2.5,3.5) node[pos=.5] {\texttt{a()}};
      \draw[visible on=<5->,fill=gray!50] (0,1) rectangle (2.5,3.5) node[pos=.5] {\texttt{a()}};
      \draw[visible on=<5->,fill=white] (0,3.5) rectangle (2.5,4.5) node[pos=.5] (b-program2) {\texttt{b()}};
      \draw[visible on=<5->] node[devil,minimum size=.5cm,anchor=west] at (b-program2.east) {};
      \draw[visible on=<5->] node[devil,mirrored,minimum size=.5cm,anchor=east] at (b-program2.west) {};

      % A'S STACK POINTER
      \begin{scope}[visible on=<5->]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (2.9,3.5) -- (2.9,6) node[black,pos=0.4] (d) {};
        \begin{scope}
          \clip (2.5,3.4) rectangle (3.5,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.4,mirror},yshift=0pt] (2.9,3.5) -- (2.9,6) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwx}} (d);
      \end{scope}

      % A'S OLD STACK POINTER
      \begin{scope}[visible on=<6->]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt,draw=white] (4.2,2.5) -- (4.2,6) node[black,pos=0.4] (d) {};
        \begin{scope}
          \clip (3.5,2.4) rectangle (4.5,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.4,mirror},yshift=0pt] (4.2,2.5) -- (4.2,6) {};
        \end{scope}
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwx}} (d);
      \end{scope}
      \pattern[visible on=<6->,pattern=north east lines, pattern color=red] (0,2.5) rectangle (2.5,3.5) node[pos=.5] {};

    \end{scope} %shift=(shifter2)
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Breaking well-bracketed control flow}
    \centering
  \begin{tikzpicture}
    % recurrent parts
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (rstk) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_\stk$};
      \coordinate (r2) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_2$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \draw[fill=gray!50] (0,-.2) rectangle (2.5,1) node[pos=.5] {\scriptsize Lower stack frames};
      \end{scope}
      % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
      \node at (1.25,6) {Call stack};
      \draw[visible on=<1->,fill=white] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      % B FIRST RETURN POINTER
      \begin{scope}[visible on=<2->]
          \clip (-.1,-1.4) rectangle (2.6,-.6);
          \draw (0,-1.5) rectangle (2.5,-.5) node[pos=.5] {\scriptsize 1st return in \texttt{a()}};
      \end{scope}
      \begin{scope}[visible on=<2>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,mirror},yshift=0pt] (2.6,-1.4) -- (2.6,-.6) node[black,pos=0.5] (d) {};
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[left] {\scriptsize\textsc{rx}} (d);
      \end{scope}

      % B FIRST RETURN POINTER
      \begin{scope}[visible on=<3-5>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,mirror},yshift=0pt] (2.6,-1.4) -- (2.6,-.6) node[black,pos=0.5] (d) {};
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[left] (label) {} (d);
        \draw (label.north west) pic {lock};
      \end{scope}

      \draw[visible on=<4-5>,fill=white] (0,2.5) rectangle (2.5,4) node[pos=.5] (b-program) {\texttt{b()}};
      \draw[visible on=<4-5>] node[devil,minimum size=.5cm,anchor=west] at (b-program.east) {};
      \draw[visible on=<4-5>] node[devil,mirrored,minimum size=.5cm,anchor=east] at (b-program.west) {};
      \draw[visible on=<4-5>,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      % B STORED RETURN POINTER
      \begin{scope}[visible on=<5-8>]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,mirror},yshift=0pt] (4.5,-1.4) -- (4.5,-.6) node[black,pos=0.5] (d) {};
        \draw[capa] (7,-.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] (label) {} (d);
        \draw (label) pic {lock};
      \end{scope}

      % B SECOND RETURN POINTER
      \begin{scope}[visible on=<7->]
        \begin{scope}
          \clip (2.9,4.6) rectangle (5.6,5.4);
          \draw (3,4.5) rectangle (5.5,5.5) node[align=center,pos=.5] {\scriptsize 2nd return in \texttt{a()}};
        \end{scope}
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,mirror},yshift=0pt] (5.6,4.6) -- (5.6,5.4) node[black,pos=0.5] (d) {};
        \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[left] (label) {} (d);
        \draw (label.west) pic {lock};
      \end{scope}

      \draw[visible on=<8->,fill=white] (0,2.5) rectangle (2.5,4) node[pos=.5] (b-program2) {\texttt{b()}};
      \draw[visible on=<8->] node[devil,minimum size=.5cm,anchor=west] at (b-program2.east) {};
      \draw[visible on=<8->] node[devil,mirrored,minimum size=.5cm,anchor=east] at (b-program2.west) {};
      \draw[visible on=<8->,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      % B LOADED STORED RETURN POINTER
      \begin{scope}[visible on=<9->]
        \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,mirror},yshift=0pt] (2.6,-1.4) -- (2.6,-.6) node[black,pos=0.5] (d) {};
        \draw[capa] (r3) edge[-latex,postaction={decorate},out=180,in=0] node[above] (label) {} (d);
        \draw (label.north) pic {lock};
      \end{scope}

      \draw[visible on=<10->,draw=red,pattern=north east lines, pattern color=red] (0,-1.4) rectangle (2.5,-.6) node[pos=.5,fill=white,inner sep=2pt] {\scriptsize 1st return in \texttt{a()}};

    \end{scope} %shift=(shifter2)
  \end{tikzpicture}
\end{frame}
% \begin{frame}
%   \frametitle{The high-level issues}
%   \begin{itemize}
%   \item
%   \end{itemize}
% \end{frame}

%\againframe<2>{title-hl}
% add a slide about formal reasoning.
% \begin{frame}
%   \frametitle{Formal reasoning}
%   % Gives absolute certainty that the calling convention works.
%   % What it takes: create mathematical model. Reason.
% \end{frame}

\begin{frame}
  \frametitle{The thesis}
  \onslide<2->{  \large \textit{Well-bracketed control flow} and \textit{local-state encapsulation} can be provably enforced on capability machines.}
\end{frame}

\begin{frame}<1-6>[label=publication-overview]
  \frametitle{Publication overview}
  \note{Mention collaborators and publication venues}
  \begin{enumerate}
  \item \textit{Reasoning About a Machine with \alert<3>{Local Capabilities} - Provably Safe Stack and Return Pointer Management}
    \begin{itemize}
    \item<2-> Calling convention
    \item<5-> Correctness proofs of examples
    \end{itemize}
  \item \alert<8>{\textit{\textsc{StkTokens}: Enforcing Well-bracketed Control Flow and Stack Encapsulation Using \alert<4>{Linear Capabilities}}}
               \note{talk about local capability enforcement and linear cap proofs}
    \begin{itemize}
    \item<2-> Calling convention
    \item<6-> Correctness proof for all programs
    \end{itemize}
  \end{enumerate}               
\end{frame}
\begin{frame}
  \frametitle{Local capabilities}
    \Large
    \textit{Register-only capabilities\onslide<2>{, essentially}}
\end{frame}

\begin{frame}
  \frametitle{Local capabilities by example}
   % A special kind of capability that can only be stored to memory by capabilities with a special write-local permission.
  \begin{tikzpicture}
    % recurrent parts
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (r1) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_1$};
      \coordinate (r2) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_2$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \foreach \x in {.25,.5,...,5}
        {
          \draw (0,\x) rectangle (2.5,\x+.25);
        };
      \end{scope}
      \node at (1.25,6) {Memory};
    % NORMAL CAPABILITY
    \begin{scope}[visible on=<1-2>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,3.5) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r1) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % LOCAL CAPABILITY
    \begin{scope}[visible on=<3-11>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,3.5) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r1) edge[-latex,dashed,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % NORMAL WRITE CAPABILITY
    \begin{scope}[visible on=<5-7>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,1) -- (2.9,2.75) node[pos=0.5] (d) {};
      \draw[capa] (r2) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rw}} (d);
    \end{scope}

    \node[visible on=<6-7>] (store) at ([xshift=-2cm]reg-file.north west) {\texttt{store}\;$r_2$\;$r_1$};
    \node[visible on=<7>] at (store.north east) {\Sadey[1.5][red]};

    % WRITE LOCAL CAPABILITY
    \begin{scope}[visible on=<9-12>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,1) -- (2.9,2.75) node[pos=0.5] (d) {};
      \draw[capa] (r2) edge[-latex,dashed,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rwl}} (d);
    \end{scope}

    \node[visible on=<10-11>] (store) at ([xshift=-2cm]reg-file.north west) {\texttt{store}\;$r_2$\;$r_1$};
    \node[visible on=<11>] at (store.north east) {\Smiley[1.5][green]};

    % LOCAL CAPABILITY
    \begin{scope}[visible on=<11>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,3.5) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (2.25,2.625) edge[-latex,dashed,postaction={decorate},out=0,in=0] node[below right] {\scriptsize\textsc{r}} (d);
    \end{scope}

    \end{scope}

  \end{tikzpicture}
  \begin{overlayarea}{\textwidth}{1em}
    \only<2-3>{Capabilities can be turned into local capabilities}
    \only<4-7>{Cannot be stored to memory}
    \only<8-11>{New write-local permission}
    \only<12>{Write-local capability must be local}
  \end{overlayarea}
\end{frame}

\begin{frame}
  \frametitle{The calling convention}
  \begin{itemize}
  \item Write-local capability for the stack
  \item Stack allocated restoration record
  \item Return capabilities are ``locked'' stack capabilities for the restoration record
  \item Revoke old stack capabilities and return capabilities by clearing unused stack and registers
  \item ... and a few other things
  \end{itemize}
   % Explain some of the calling convention
\end{frame}

\begin{frame}
  \frametitle{The calling convention by example}
  \begin{tikzpicture}
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (rstk) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_\stk$};
      \coordinate (rret) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_\ret$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \draw[fill=gray!50] (0,-.2) rectangle (2.5,1) node[pos=.5] {\scriptsize Lower stack frames};
      \end{scope}
      \node at (1.25,6) {\footnotesize Call stack in memory};
      \begin{scope}[visible on=<1-6>]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,dashed,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{rwlx}} (e);
      \end{scope}

      \draw[visible on=<2-8>] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      \begin{scope}[visible on=<3-8>]
        \draw (0,2.5) rectangle (2.5,3) node[pos=.5] {\scriptsize Restoration record};
        \begin{scope}[visible on=<4>]
          \filldraw[shading=axis,left color=black!15,right color=white,shading angle=90] (2.5,3) -- (4,3.5) -- (4,1.25) -- (2.5,2.5) -- cycle;
          \draw[fill=white] (4,1.25) rectangle (7,3.5);
          \draw (4,2) rectangle (7,3.5) node[pos=.5] {\scriptsize Restoration instructions};
          \draw (4,1.625) rectangle (7,2) node[pos=.5] {\scriptsize Return capability};
          \draw (4,1.25) rectangle (7,1.625) node[pos=.5] {\scriptsize Stack capability};
        \end{scope}

      \end{scope}

      \begin{scope}[visible on=<6->]
        \begin{scope}
          \clip (2.5,.9) rectangle (3.5,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.35,mirror},yshift=0pt] (3.1,1) -- (3.1,6) node[black,pos=0.35] (e) {};
        \end{scope}
        \draw[capa] (rret) edge[-latex,dashed,postaction={decorate},out=135,in=0,pos=.75] node[above] (lab) {} (e);
        \draw (lab) pic {lock};
      \end{scope}

      \begin{scope}[visible on=<7->]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,3) -- (2.6,6) node[black,pos=0.25] (astk) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,dashed,postaction={decorate},out=180,in=0] node[above] (astk-label) {\scriptsize\textsc{rwlx}} (astk);
      \end{scope}

      \begin{scope}[visible on=<8->]
        \foreach \x in {3,3.25,...,5}{
          \draw (0,\x) rectangle (2.5,\x+.25) node[pos=.5] {\scriptsize 0};
        }
        \node at (r3) {0};
        \node at (r4) {0};
      \end{scope}

      \draw[visible on=<9->,fill=white] (0,3) rectangle (2.5,4.5) node[pos=.5] (a-program) {\texttt{b()}};
      \draw[visible on=<9->,fill=gray!50] (0,2.5) rectangle (2.5,3) node[pos=.5] {\scriptsize Restoration record};
      \draw[visible on=<9->,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      \draw[visible on=<10->] node[devil,minimum size=.5cm,anchor=west] at (a-program.east) {};
      \draw[visible on=<10->] node[devil,minimum size=.5cm,mirrored,anchor=east] at (a-program.west) {};
      \draw[visible on=<11>] node[circle, fit=(rstk),draw=red,thick,inner sep=12pt] {};
      \draw[visible on=<12>] node[circle, fit=(rret),draw=red,thick,inner sep=12pt] {};
    \end{scope}
  \end{tikzpicture}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Formal result}
    \begin{lemma}[Example correctness]
No matter what program context the program
    \begin{lstlisting}[basicstyle=\small\ttfamily,escapeinside=||,keywords={fun, let, in, ref, assert}]
    fun _ => 
      let x = ref 0 in
        fun callback =>
          x := 0;
          callback();
          x := 1;
          callback();
          assert(x == 1)
    \end{lstlisting}
\vspace{-1em}
is used in, the assertion never fails.
  \end{lemma}
\end{frame}

\begin{frame}
  \frametitle{Summary of the calling convention}
  \huge It's great!\\
    \begin{flushright}
    \onslide<2->{\normalsize ...except for all the stack clearing.}
  \end{flushright}
  % Large amount of memory clearing necessary.
  % Not immediately obvious whether this is feasible.
\end{frame}

\againframe<8>{publication-overview}

\begin{frame}
  \frametitle{Linear capabilities}
  % Briefly describe linear capabilities and stktokens
  \begin{quote}
    \large
    Capabilities that cannot be duplicated
  \end{quote}
\end{frame}

\begin{frame}
  \frametitle{Linear capabilities by example}
  \begin{tikzpicture}
    % recurrent parts
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (r1) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_1$};
      \coordinate (r2) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_2$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \foreach \x in {.25,.5,...,5}
        {
          \draw (0,\x) rectangle (2.5,\x+.25);
        };
      \end{scope}
      \node at (1.25,6) {Memory};
    \node[visible on=<1-6>] (move) at ([xshift=-2cm]reg-file.north west) {\texttt{move}\;$r_3$\;$r_1$};
    % NORMAL CAPABILITY
    \begin{scope}[visible on=<1-2>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r1) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % NORMAL CAPABILITY MOVED
    \begin{scope}[visible on=<2>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r3) edge[-latex,postaction={decorate},out=180,in=0] node[below] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % LINEAR CAPABILITY
    \begin{scope}[visible on=<3-4>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r1) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % LINEAR CAPABILITY MOVED
    \begin{scope}[visible on=<5-6>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r3) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % CLEARED R1
    \node[visible on=<5->] at (r1) {0};

    % LINEAR CAPABILITY SPLIT
    \begin{scope}[visible on=<7-8>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,3) node[pos=0.5] (d) {};
      \draw[capa] (r3) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,3) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r2) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % LINEAR CAPABILITY SPLICED
    \begin{scope}[visible on=<9-10>]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r2) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    % CLEARED R3
    \node[visible on=<9-10>] at (r3) {0};

    % LINEAR CAPABILITY SPLIT
    \begin{scope}[visible on=<11->]
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,2) -- (2.9,2.75) node[pos=0.5] (d) {};
      \draw[capa] (r3) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
      \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (2.9,3.25) -- (2.9,5) node[pos=0.5] (d) {};
      \draw[capa] (r2) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{r}} (d);
    \end{scope}

    \node[visible on=<12->] at (2.9,3) {\Sadey[1.5][red]};
    \end{scope}

  \end{tikzpicture}
  \begin{overlayarea}{\textwidth}{1em}
    \only<4-5>{Moving a linear capability clears the source}
    \only<6-9>{Linear capabilities can be split}\only<8-9>{ and spliced}
    \only<10->{Splice fails for non-adjacent capabilities}
  \end{overlayarea}
\end{frame}

\begin{frame}
  \frametitle{\stktokens{}}
  \begin{itemize}
  \item Use a \textit{unique} linear capability for the stack
  \item Require that callees return their stack capability
  \item ... and a few other things
  \end{itemize}  
\end{frame}

\begin{frame}
  \frametitle{\stktokens{} by example}
  \begin{tikzpicture}
    \begin{scope}
      \draw (0.5,2.5) node (reg-file) {Registers};
      \begin{scope}
        \foreach \x in {-2,-1,...,1}
        {
          \draw[fill=white] (0,\x) rectangle (1,\x+1) node[pos=.5,color=mygreen] {};
        };
      \end{scope}
      \coordinate (rstk) at (.5,1.5) {};
      \node[anchor=west] at (1.1,1.5) {$r_\stk$};
      \coordinate (rret) at (.5,.5) {};
      \node[anchor=west] at (1.1,.5) {$r_\ret$};
      \coordinate (r3) at (.5,-.5) {};
      \node[anchor=west] at (1.1,-.5) {$r_3$};
      \coordinate (r4) at (.5,-1.5) {};
      \node[anchor=west] at (1.1,-1.5) {$r_4$};
    \end{scope}

    \coordinate (shifter2) at (-8,-2.25);
    \begin{scope}[shift=(shifter2)]
      \begin{scope}
        \clip (-.1,-.1) rectangle (2.6,5.6);
        \fill[fill=white] (0,0) rectangle (2.5,5.5);
        \draw (0,0) -- (0,5.5);
        \draw (2.5,0) -- (2.5,5.5);
        \draw[fill=gray!50] (0,-.2) rectangle (2.5,1) node[pos=.5] {\scriptsize Lower stack frames};
      \end{scope}
      \node at (1.25,6) {\footnotesize Call stack in memory};
      \draw[visible on=<1-2>] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};

      % STACK CAPABILITY
      \begin{scope}[visible on=<1>]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rw}} (e);
      \end{scope}

      % SPLIT STACK CAPABILITY
      \begin{scope}[visible on=<2-4>]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.10,mirror},yshift=0pt] (2.6,2.5) -- (2.6,6) node[black,pos=0.10] (e) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rw}} (e);

        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.80,mirror},yshift=0pt] (2.6,1) -- (2.6,2.5) node[black,pos=0.80] (e) {};
        \end{scope}
        \draw[capa] (rret) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[below] (ret-label) {} (e);
        \draw[visible on=<2-3>] (ret-label.south) pic {lock};
        \node[visible on=<4>] at (ret-label) {\scriptsize\textsc{rw}};
      \end{scope}

      % CALL A
      \draw[visible on=<3>,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};
      \draw[visible on=<3>,fill=white] (0,2.5) rectangle (2.5,4) node[pos=.5] {\texttt{b()}};

      % RETURN TO B
      \draw[visible on=<4-5>] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};
      \begin{scope}[visible on=<5>]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.25,mirror},yshift=0pt] (2.6,1) -- (2.6,6) node[black,pos=0.25] (e) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rw}} (e);
      \end{scope}

      % SAY WE CALLED A...
      \begin{scope}[visible on=<6-7>]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.10,mirror},yshift=0pt] (2.6,2.5) -- (2.6,6) node[black,pos=0.10] (e) {};
        \end{scope}
        \draw[capa] (rstk) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above] {\scriptsize\textsc{rw}} (e);
      \end{scope}
      \begin{scope}[visible on=<6->]
        \begin{scope}
          \clip (2.5,.9) rectangle (3,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.80,mirror},yshift=0pt] (2.6,1) -- (2.6,2.5) node[black,pos=0.80] (e) {};
        \end{scope}
        \draw[capa] (rret) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[below] (ret-label) {} (e);
        \draw[visible on=<6-8>] (ret-label.south) pic {lock};
        \node[visible on=<9->] at (ret-label) {\scriptsize\textsc{rw}};
      \end{scope}

      \draw[visible on=<6-8>,fill=gray!50] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};
      \draw[visible on=<6-8>,fill=white] (0,2.5) rectangle (2.5,4) node[pos=.5] (a-frame) {\texttt{b()}};

      % ... AND A IS EVIL
      \draw[visible on=<7-8>] node[devil,minimum size=.5cm,mirrored,anchor=east] at (a-frame.west) {};
      \draw[visible on=<7-8>] node[devil,minimum size=.5cm,anchor=west] at (a-frame.east) {};

      \begin{scope}[visible on=<8>]
        \begin{scope}
          \clip (2.5,.9) rectangle (4,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.10,mirror},yshift=0pt] (3.6,2.5) -- (3.6,6) node[black,pos=0.10] (e) {};
        \end{scope}
        \draw[capa] ([xshift=-.5cm]reg-file.north west) edge[-latex,densely dotted,postaction={decorate},out=180,in=0] node[above left] {\scriptsize\textsc{rw}} (e);
      \end{scope}
      \node[visible on=<8->] at (rstk) {0};

      \draw[visible on=<9->] (0,1) rectangle (2.5,2.5) node[pos=.5] {\texttt{a()}};
      \begin{scope}[visible on=<9->]
        \begin{scope}
          \clip (2.5,.9) rectangle (4,5.75);
          \draw[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.10,mirror},yshift=0pt,draw=black!50] (3.6,2.5) -- (3.6,6) node[black,pos=0.10] (e) {};
        \end{scope}
        \draw[decoration={markings,mark=at position 0 with {\fill[black!50] circle(1.5pt);}}] ([xshift=-.5cm]reg-file.north west) edge[-latex,densely dotted,postaction={decorate},out=180,in=0,draw=black!50] node[above left] {\color{black!50}\scriptsize\textsc{rw}} (e);
      \end{scope}
      \draw[visible on=<10->] (3.1,2.5) node {\Sadey[1.5][red]};
    \end{scope}
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{The formal result}
  \begin{theorem}
  \stktokens{} enforces \textit{local-state encapsulation} and \textit{well-bracketed control flow} for all programs.
\end{theorem}
\begin{proof}
\begin{center}
    \Large
By fully-abstract overlay semantics.
%    \tikz[remember picture] \node[inner sep=0pt] (fal) {};fully-abstract\tikz[remember picture] \node[inner sep=0pt] (far) {}; \tikz[remember picture] \node[inner sep=0pt] (osl) {};overlay semantics\tikz[remember picture] \node[inner sep=0pt] (osr) {};
  \end{center}
\end{proof}
  \note{Why something new? No good definition of WBCF and LSE}
%   \begin{tikzpicture}[remember picture, overlay]
%     \coordinate (fall) at ([xshift=.2cm,yshift=0.3cm]fal);
%     \coordinate (farr) at ([xshift=-0.2cm,yshift=0.1cm]far);

%     \coordinate (osll) at ([xshift=0.2cm,yshift=0.3cm]osl);
%     \coordinate (osrr) at ([xshift=-0.2cm,yshift=0.1cm]osr);

%     \draw<2>  node[fit=(fall) (farr), thick, rectangle, draw=red, rounded corners, inner sep=12pt] {};
%     \draw<3>  node[fit=(osll) (osrr), thick, rectangle, draw=red, rounded corners, inner sep=12pt] {};

%   \end{tikzpicture}
 \end{frame}

\newsavebox{\syntaxplain}
\begin{lrbox}{\syntaxplain}
\begin{minipage}{2.5cm}
\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
    move  rtmp1 42
   !\tikz[remember picture] \node[] (calltl) {};!store rstk rtmp1
    cca   rstk -1
    geta  rtmp1 rstk
    cca   rretc 5
    move  rtmp1 pc
    cca   rtmp1 -20
\end{lstlisting}
\end{minipage}
\begin{minipage}{2.5cm}
\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
    load  rtmp1 rtmp1
    cca   rtmp1 -21
    cseal rretd rtmp1
    move  rretc pc
    xjmp  r1 r2
    cseal rretc rtmp1
    move  rtmp1 0
  \end{lstlisting}
\end{minipage}
\begin{tikzpicture}[remember picture, overlay]
  \draw node[rectangle, fit=(calltl),fill=red,fill opacity=0, rounded corners, inner sep=2pt] (call) {};
\end{tikzpicture}
\end{lrbox}

\newsavebox{\syntax}
\begin{lrbox}{\syntax}
\begin{minipage}{2.5cm}
\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
    move  rtmp1 42
   !\tikz[remember picture] \node[] (calltl) {};!store rstk rtmp1
    cca   rstk -1
    geta  rtmp1 rstk
    cca   rretc 5   !\tikz[remember picture] \node[] (callbr) {};!
    move  rtmp1 pc
    cca   rtmp1 -20
\end{lstlisting}
\end{minipage}
\begin{minipage}{2.5cm}
\begin{lstlisting}[basicstyle=\tiny\ttfamily,escapechar=!]
    load  rtmp1 rtmp1
    cca   rtmp1 -21
    cseal rretd rtmp1
    move  rretc pc
   !\tikz[remember picture] \node[] (retl) {};!xjmp  r1 r2 !\tikz[remember picture] \node[] (retr) {};!
    cseal rretc rtmp1
    move  rtmp1 0
  \end{lstlisting}
\end{minipage}
\begin{tikzpicture}[remember picture, overlay]
  \draw node[rectangle, fit=(retl) (retr),fill=red,fill opacity=0.6, rounded corners,inner sep=3pt] (ret) {};
  \draw node[rectangle, fit=(retl) (retr)] [yshift=-.1cm] (rettext) {\footnotesize\textbf{\texttt{return}}};

  \draw node[rectangle, fit=(calltl) (callbr),fill=red,fill opacity=0.6, rounded corners, inner sep=2.5pt] (call) {};
  \draw node[rectangle, fit=(calltl) (callbr)] (calltext) [yshift=-0.1cm] {\textbf{\texttt{call}}};
\end{tikzpicture}

\end{lrbox}

\begin{frame}[fragile]
  \frametitle{Overlay semantics}

\begin{tikzpicture}
  \node<1-3> (syntax) {\usebox{\syntaxplain}};
  \node[anchor=south west] (syntax-text) at ([yshift=-.5cm]syntax.north west) {\emph{Syntax}};
  \node[anchor=west] (capmac) at ([xshift=1.1cm]syntax.east) {Capability machine};
  \node[visible on=<1-3>,anchor=north] (gears) at ([yshift=-1cm]syntax.south) {\includegraphics[width=2cm]{gears2.png}};
  \node[anchor=west] (semantics-text) at ([yshift=-1.5cm]syntax.west) {\emph{Semantics}};
  \node[visible on=<2->,anchor=west] at ([xshift=-.5cm,yshift=-1pt]capmac.south) {\color{red}with a stack!};
  \node[visible on=<4->,anchor=north] (gears) at ([yshift=-1cm]syntax.south) {\includegraphics[width=2cm]{gears3.png}};
  \node<4-> (syntax) {\usebox{\syntax}};
  \begin{scope}[visible on=<3->]
  \node[anchor=west] (plus) at ([xshift=.5cm]gears.east) {\large +};
  \coordinate (stack) at ([xshift=2.2cm]gears.south east);
  \coordinate (stack-top-right) at ($(stack) + (2.5,2)$);
  \begin{pgfonlayer}{bg}
    \draw<3-> node[rectangle, fit=(plus) (stack) (stack-top-right), fill=red, fill opacity=0.6, rounded corners] {}; 
  \end{pgfonlayer}
  \begin{scope}[shift=(stack)]
    \clip (-.1,0) rectangle (3,3);
    \draw (0,-.1) rectangle (2.5,2);
    \draw (0,-.1) rectangle (2.5,.5) node[pos=.5] {\scriptsize ...};
    \draw (0,.5) rectangle (2.5,1) node[pos=.5] {\scriptsize\texttt{a()}};
    \draw (0,1) rectangle (2.5,1.5) node[pos=.5] {\scriptsize\texttt{b()}};
    \draw (0,1.5) rectangle (2.5,2) node[pos=.5] {\scriptsize\texttt{a()}};
  \end{scope}
  \end{scope}
  \node at ($(stack.south)+(0,-0.2)$) {};
\end{tikzpicture}

\end{frame}

\begin{frame}
  \frametitle{Fully-abstract overlay semantics}
  \begin{center}
  \begin{tikzpicture}
    \draw[visible on=<1->] node (capmac) {Capability machine};

    \draw[visible on=<2->] node[above=of capmac ,yshift=1cm] (overlay) {Capability machine};
    \draw[visible on=<2->] node[anchor=west] at ([xshift=-.5cm,yshift=-1pt]overlay.south) {\color{red}with a stack!};
    \draw[visible on=<2->] (-2.5,1.4) edge[dashed] (2.5,1.4);

    \draw[visible on=<3->] ([xshift=5pt]overlay.east) edge[-latex, bend left,thick] node[right] (trans-label) {\footnotesize Identity translation} ([xshift=5pt]capmac.east);

    \begin{scope}[visible on=<4->]
      \draw node (fu) at ([xshift=.5cm,yshift=-1.2cm]trans-label.south east) {\scriptsize Fully abstract};
      \draw (trans-label) -- (fu);
    \end{scope}
  \end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}
  \frametitle{Summary of \stktokens{}}
  \huge It's great!\\
    \begin{flushright}
      \onslide<2->{\normalsize ...except for the linear capabilities}
    \end{flushright}
\end{frame}

\begin{frame}
  \frametitle{Conclusions from PhD dissertation}
  \begin{itemize}
  \item \textit{Well-bracketed control flow} and \textit{local-state encapsulation} can be enforced on capability machines.
  \item We can prove the enforcement correct using advanced reasoning techniques.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Looking forward...}
  \begin{itemize}
  \item This dissertation is a step on the way towards \emph{real} secure compilers that target capability machines. \onslide<2->{We still need to:}
    \begin{itemize}
      \item<2-> Enforce other high-level abstractions.
      \item<3-> Mechanise all proofs.
        \begin{itemize}
          \item<4-> Program logic?
          \item<4-> Proof automation?
      \end{itemize}
      \item<5-> Proof of concept secure compiler.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \centering

  \huge Thank you!
\end{frame}
\appendix
\begin{frame}
  \begin{center}
    \begin{quote}
      \Huge Bonus slides!
    \end{quote}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{The mistake}
  For a register file with the contents
  \begin{itemize}
  \item $((\textsc{e},\textsc{global}),b,e,a)$ in register $r_1$
  \item $\mathit{encode}((\textsc{e},\textsc{local}))$ in register $r_2$
  \end{itemize}
  executing $\texttt{restrict}\;r_1\;r_2$ would result in a register file with
  \begin{itemize}
  \item $((\textsc{e},\textsc{local}),b,e,a)$ in register $r_1$
  \end{itemize}
  seemingly fine\onslide<2>{, but the logical relation does not account for this.}
\end{frame}

\begin{frame}
  \frametitle{Fixing the mistake}
  Either
  \begin{itemize}
  \item Prohibit \texttt{restrict} from restricting an enter-capability from global to local; or
  \item Fix the logical relation:
    Insert $\forall g' \sqsubseteq g$ where appropriate, e.g. in enter-condition:
  \end{itemize}
  \vspace{1em}
  \includegraphics[width=\textwidth]{econd}
\end{frame}

\begin{frame}
  \frametitle{Flavours of linear capabilities}
  \begin{columns}
\begin{column}{.7\textwidth}
  \begin{itemize}
  \item unique linear capability\\
    \begin{itemize}
    \item No alias
    \item \stktokens{}-paper
  \end{itemize}
\item linear\\
  \begin{itemize}
  \item May have aliases
  \item CHERI
  \end{itemize}
  \end{itemize}
\end{column}
\begin{column}{.3\textwidth}
  \begin{tikzpicture}
    \node (ul) {unique linear};
    \node[below=of ul] (n) {normal};
    \node[below=of n] (l) {linear};
    \draw (ul) -- (n) -- (l);
  \end{tikzpicture}
\end{column}
\end{columns}
\end{frame}

\end{document}

% \begin{frame}<1-2>[label=machine]
% %  \frametitle{Machine}
%   \begin{columns}
%     \begin{column}{.6\textwidth}
%       \begin{tikzpicture}[scale=.45, every node={scale=.5}]
%         \scope
%           \clip (-.1,-.1) rectangle (6.1,14.9);
%           \fill[draw,fill=white] (0,0) rectangle (6,15);
%         \endscope
%         % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward}
%         % (-2,15);
%         \foreach \X in {0,0.5,...,14} {
%           \fill[draw,fill=white] (0,\X) rectangle (6,\X+.5) node[pos=.5] { };
%         };
%         \draw (-0.6,-0.1) node[rotate=45, align=right] { \scriptsize\texttt{0x00} };
%         \draw (-0.6,3.9)  node[rotate=45, align=right] { \scriptsize\texttt{0x08} };
%         \draw (-0.6,7.9)  node[rotate=45, align=right] { \scriptsize\texttt{0x10} };
%         \draw (-0.6,11.9) node[rotate=45, align=right] { \scriptsize\texttt{0x16} };

%   \begin{scope}
%     \clip (9.9,2.6) rectangle (11.6,11.6);
%     \foreach \X in {0,1.5,...,7.5} {
%       \fill[draw,fill=white] (10,10 - \X) rectangle (11.5,11.5-\X) node[pos=.5] { };
%     };
%   \end{scope}
%   \draw (10.75,12.25) node[align=center] {Register File};
%   \draw (6.5,14.325) node[anchor=west, align=left] {Memory};

%   % Mem data
%   \draw<1-> (3,0.75) node { \scriptsize\texttt{0x2a}};
%   \draw<1-> (3,1.25) node { \scriptsize\texttt{0x02}};
%   \draw<1-> (3,1.75) node { \scriptsize\texttt{0x0e}};
%   \draw<1-> (3,2.75) node { \scriptsize\texttt{0x0a}};
%   \draw<1-> (3,2.25) node { \scriptsize\texttt{0x13}};
%   \draw<1-> (3,3.75) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,3.25) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,4.25) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,4.75) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,5.75) node { \scriptsize\texttt{0x03}};
%   \draw<1-> (3,6.25) node { \scriptsize\texttt{0x14}};
%   \draw<1-> (3,6.75) node { \scriptsize\texttt{0x15}};
%   \draw<1-> (3,7.25) node { \scriptsize\texttt{0x0b}};
%   \draw<1-> (3,7.75) node { \scriptsize\texttt{0x0a}};
%   \draw<1-> (3,8.25) node { \scriptsize\texttt{0x0f}};
%   \draw<1-> (3,8.75) node { \scriptsize\texttt{0x13}};
%   \draw<1-> (3,9.25) node { \scriptsize\texttt{0x1b}};
%   \draw<1-> (3,9.75) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,10.25) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,10.75) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,11.25) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,11.75) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,12.75) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,13.25) node { \scriptsize\texttt{0x00}};
%   \draw<1-> (3,13.75) node { \scriptsize\texttt{0x40}};
%   \draw<1-> (3,14.25) node { \scriptsize\texttt{0x05}};
%   % Reg data
%   \draw<1-> (10.75,10.75) node { \scriptsize\texttt{0x08}};
%   \draw<1-> (10.75,7.75) node { \scriptsize\texttt{0x18}};
%   \draw<1-> (10.75,6.25) node { \scriptsize\texttt{0x10}};
%   \draw<1-> (10.75,4.75) node { \scriptsize\texttt{0x02}};

%   % Mem pointers
%   \draw<1> (3,12.25) node { \scriptsize\texttt{0x10}};
%   \draw<1> (3,5.25) node { \scriptsize\texttt{0x0b}};
%   \draw<1> (3,0.25) node { \scriptsize\texttt{0x08}};
%   % Reg pointers
%   \draw<1> (10.75,9.25) node { \scriptsize\texttt{0x00}};

%   % Mem pointers HL
%   \draw<2> (3,12.25) node { \color{red} \scriptsize\texttt{0x10}};
%   \draw<2> (3,5.25) node { \color{red} \scriptsize\texttt{0x0b}};
%   \draw<2> (3,0.25) node { \color{red} \scriptsize\texttt{0x08}};
%   % Reg pointers HL
%   \draw<2> (10.75,9.25) node { \color{red} \scriptsize\texttt{0x00}};

%   \draw<3> (10.75,9.25) node { };

%   % Mem capabilities
%   \draw<4->[capa] (3,12.25) edge[postaction={decorate}] (6.1,12.25 );
%   \draw<4> (6.1,12.25) edge[-latex,in=0,out=0] (6.1,8.25);
%   \draw<4->[capa] (3,5.25) edge[postaction={decorate}] (6.1,5.25);
%   \draw<4> (6.1,5.25) edge[-latex,in=0,out=0] (6.1,6.75);
%   \draw<4->[capa] (3,0.25) edge[postaction={decorate}] (6.1,0.25);
%   \draw<4> (6.1,0.25) edge[-latex,in=0,out=0] (6.1,4.25);

%   % Reg pointers HL
%   \draw<4->[capa] (10.75,9.25) edge[postaction={decorate}] (10.5,9.25);
%   \draw<4> (10.5 ,9.25) edge[-latex,in=0,out=180] (6.1,0.25);

%   % Mem capabilities \w auth
%   \draw<5> (6.1,12.25) edge[-latex,in=0,out=0] (6.5,8.25);
%   \draw<5->[decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.55,mirror},yshift=0pt] (6.3,5.5) -- (6.3,10.5) node (e) [black,midway,xshift=0.8cm] {};

%   \draw<5> (6.1,5.25) edge[-latex,in=0,out=0] (6.3,6.75);
%   \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.1,6) -- (6.1,7.5) node (e) [black,midway,xshift=0.8cm] {};

%   \draw<5> (6.1,0.25) edge[-latex,in=0,out=0] (6.5,4.25);
%   \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.5,mirror},yshift=0pt] (6.3,3) -- (6.3,5.5) node (e) [black,midway,xshift=0.8cm] {};

%   % Reg pointers HL \w auth
%   \draw<5> (10.5,9.25) edge[-latex,in=0,out=180] (6.3,0.355);
%   \draw<5-> [decorate,decoration={brace,amplitude=2pt,raise=0pt,aspect=0.1,mirror},yshift=0pt] (6.1,0) -- (6.1,3.5) node (e) [black,midway,xshift=0.8cm] {};

%   % Mem capabilities \w auth & perm
%   \draw<6-> (6.1,12.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{r}} (6.5,8.25);

%   \draw<6-> (6.1,5.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{rw}} (6.3,6.75);

%   \draw<6-> (6.1,0.25) edge[-latex,in=0,out=0] node[right] {\scriptsize \textsc{r}} (6.5,4.25);

%   % Reg pointers HL \w auth & perm
%   \draw<6-> (10.5,9.25) edge[-latex,in=0,out=180] node[right] {\scriptsize \textsc{rx}} (6.3,0.355);
% \end{tikzpicture}
% \end{column}
% \begin{column}{0.4\textwidth}
%   \onslide<3->{Replace pointers with capabilities}
%   \begin{itemize}
% nnn  \item<4-> Pointer
%   \item<5-> Range of authority
%   \item<6-> Permission
%   \end{itemize}
%   \onslide<7->{Dynamically enforced}
%   \begin{itemize}
%   \item<8-> Cap. aware instructions
%   \item<9-> Tagged memory
%   \end{itemize}
% \end{column}
% \end{columns}
% \end{frame}

### 2020-03-11 popl-2019-journal-v/discussion.tex
Full pass over paper, ready to submit?

@@
 \begin{jversion}
-  A compiler is secure when it enforces the properties of high-level languages which begs the question what properties should we enforce.
+  A compiler is secure when it enforces the properties of high-level languages which begs the question what properties we should enforce.
   When it comes to fully-abstract compilation, then the answer is that all the properties of the high-level language should be enforced, so the real question is what high-level language we would like.
   \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/CC, then we need to come up with a different enforcement scheme.
-  Further, many high-level languages have exceptions which is yet another form of control-flow which is also not supported by \stktokens{}.
+  Further, many high-level languages have exceptions as another form of control-flow not (yet) supported by \stktokens{}.
   This goes to show how we must consider what high-level language we want in order to answer the question of what properties we must enforce to get a fully-abstract compiler.
@@
   % TODO new paragraph: Machine checked proof. MANY details, any future work should be machine checked?
-  One challenge in full-abstraction proofs is to relate the translation of program to the program it was translated from.
-  Such a relation is often expressed as a back-translation  \citep{devriese_modular_2017}, i.e.\ a translation from the target language to the source language.
+  One challenge in full-abstraction proofs is to construct source language contexts that emulate the behavior of an arbitrary target language context.
+  This construction is known as a back-translation  \citep{devriese_modular_2017}, i.e.\ a translation from the target language to the source language.
   When we use an overlay semantics, the back-translation becomes trivial because the source and target language are syntactically the same, so the identity can be used as the back-translation.

### 2020-03-11 popl-2019-journal-v/paper.tex
Full pass over paper, ready to submit?

@@
 Generally, properties that hold in the source language but not in the target language need to be somehow enforced by the compiler.
-Two properties that hold in many high-level source languages are well-bracketed control flow and encapsulation of local state, but they are not enforced after compilation to assembly.
-
-Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations.
-At the assembly level, this property does not hold immediately. 
+Two properties that hold in many high-level source languages are well-bracketed control flow and encapsulation of local state, but they are usually not enforced after compilation to assembly.
+
+Well-bracketed control flow (WBCF) expresses that invoked functions must either return to their callers, invoke other functions themselves or diverge, and generally holds in programming languages that do not offer a primitive form of continuations (or related features).
+At the assembly level, this property does not hold. 
 Invoked functions get direct access to return pointers that they are supposed to jump to a single time at the end of their execution.
 There is, however, no guarantee that untrusted assembly code respects this intended usage.
-In particular, a function may invoke return pointers from other stack frames than its own: either frames higher in the call stack or ones that no longer exist as they have already returned.
+In particular, adversarial code may invoke return pointers that were intended to be called from other stack frames than theirs: either from frames higher in the call stack or from ones that no longer exist as they have already returned.
 
 Local state encapsulation (LSE) is the guarantee that when a function invokes another function, its local variables (saved on its stack frame) will not be read or modified until the invoked function returns.
-At the assembly level, this property also does not hold immediately.
+At the assembly level, this property also does not hold.
 The calling function's local variables are stored on the stack during the invocation, and functions are not supposed to touch stack frames other than their own.
@@
 
-To enforce these properties, target language security primitives are needed that can be used to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
+To enforce these properties, target language security primitives are needed to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
 The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
 More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri_2015}.
-These processors use tagged memory to enforce a strict separation between integers and {\itshape capabilities}: pointers that carry authority.
+These processors use tagged memory to enforce a strict distinction between integers and {\itshape capabilities}: pointers that carry authority.
 Capabilities come in different flavours.
@@
 For example, on a recent capability machine called CHERI they take the form of pairs of capabilities that represent the code and data parts of the closure.
-Each of the two capabilities are sealed with a common seal which make them opaque.
-The hardware transparently unseals the pair upon invocation~\citep{watson_capability_2015,watson_fast_2016}.
+Both capabilities are sealed with a common seal.
+This makes them opaque and unusable until they are invoked.
+When they are invoked, the hardware, or a special OS-provided exception handler, transparently unseals the pair~\citep{watson_capability_2015,watson_fast_2016}.
 
@@
 % KJAA: 'Boundary crossing' is new here.
-However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearingt the entire unused part of the stack, an operation that may be prohibitively expensive.
+However, a problem with this approach is that revoking the local stack and return capabilities on every security boundary crossing requires clearing the entire unused part of the stack, an operation that may be prohibitively expensive (although the hardware could optimize the clearing to a significant extent \citep{joannou_efficient_2017}).
 
@@
 The hardware prevents these capabilities from being duplicated.
-We propose to make stack and return pointers linear and require components to hand them out in cross-component invocations and to give them back on return.
-The non-duplicability of linear capabilities together with some security checks allow us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
+We propose to make stack and return pointers linear and make components hand them out in cross-component invocations and require them back on return.
+The non-duplicability of linear capabilities together with some security checks allows us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
 
@@
   This text is an extended version of a paper that was presented at POPL 2019~\citep{skorstengaard_stktokens_2019}.
-  Compared to the earlier text, this version adds and explains many of the details that were left out in the conference version due to space restrictions.
-  The added details include a better motivation of sealing (\sectionname~\ref{sec:purpose-sealing}), well-formedness components (\sectionname~\ref{sec:well-form-reas}), and reasonable components (\sectionname~\ref{sec:reasonable-components}).
+  Compared to the earlier text, this version adds and explains aspects of our work that were left out in the conference version due to space restrictions.
+  The added details include a better motivation of sealing (\sectionname~\ref{sec:purpose-sealing}), the requirements of well-formedness (\sectionname~\ref{sec:well-form-reas}) and reasonability of components (\sectionname~\ref{sec:reasonable-components}).
   The section on proving full-abstraction (\sectionname~\ref{sec:fa-proof}) has been rewritten with details and explanations about the formal method used for the full-abstraction proof.
-  This includes a detailed description and motivation of the world (\sectionname~\ref{subsec:worlds}) and logical relation (\sectionname~\ref{subsec:logical-relation}).
-  This paper is accompanied by a technical report~\citep{technical_report_popl} with the elided details and proof.
+  This includes a description and motivation of the Kripke world (\sectionname~\ref{subsec:worlds}) and logical relation (\sectionname~\ref{subsec:logical-relation}).
+  This paper is accompanied by a technical report~\citep{technical_report_popl} with technical details and proofs.
 \end{jversion}
@@
 Any reasonable capability machine needs a way to set up boundaries between security domains with different authorities.
-It also must have a way to cross these boundaries such that (1) the security domain we move from can encapsulate and later regain its authority and (2) the security domain we move to regains all of its authority.
+It also must have a way to cross these boundaries such that (1) the security domain we move from can encapsulate itself and later regain its authority and (2) the security domain we move to regains all of its authority.
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
@@
 In order to seal a sealable with a seal $\sigma$, it is necessary to have the authority to do so.
-The permission to make sealed capabilities is represented by a sets of seals $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
-A set of seals is a capability that represents the authority to seal sealables with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In spirit of memory capabilities, a set of seals has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
+The permission to make sealed capabilities is represented by a set of seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
+Such a capability represents the authority to seal other capabilities with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
+In spirit of memory capabilities, a set of seals capability has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
 % On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
-As we will see later, sealed capabilities can be unsealed with an \texttt{xjmp}, an operation that operates on a pair of capabilities sealed with the same seal.
-The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part of the pair) available to the invoked code.
+As we will see later, sealed capabilities get unsealed when they are invoked using an \texttt{xjmp}, an operation that operates on a pair of capabilities sealed with the same seal.
+The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part) available to the invoked code.
 The combination of sealed capabilities and \texttt{xjmp} gives (1) and (2).
@@
 We define register files as functions from register names to words.
-Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ partial functions).
+Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ they are partial functions).
 $\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
@@
 Generally speaking, a capability machine needs an instruction for reducing the range of authority of a capability.
-Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two rather than reducing the range of authority (\texttt{split}).
+Because \trgcm{} has linear capabilities, the instruction for this splits the capability in two (\texttt{split}) rather than reducing the range of authority.
 The reverse is possible using \texttt{splice}.
@@
   The instructions $\tgetb{}{}$ and $\tgete{}{}$ also work on sets of seals.
-  The instruction $\tisptr{}{}$ returns an integer representation of the type of a word which allows programs to be defensive in the sense that they can check whether a word is a capability before they use it.
+  The instruction $\tisptr{}{}$ returns an integer representation of the type of a word (capability or number).
   \trgcm{} also has arithmetic instructions $\tplus{}{}{}$, $\tminus{}{}{}$, and $\tlt{}{}{}$.
@@
   % Sealing example:
-  To motivate the necessity of an encapsulation mechanism like sealing, consider the following example.
-  We, a trusted piece of code, want to transfer control to code that we distrust, and we want to give them the means to return to us.
+  To motivate the necessity of an encapsulation mechanism like sealing, consider the scenario where we are executing some trusted code and want to transfer control to untrusted code.
+  We want to give the untrusted code the means to return to us.
   That is, we need to give them a return capability.
-  If we did not have an encapsulation mechanism, our only option would be to give them an executable capability for the address we want them to return to.
-  The untrusted code could use the return capability as intended, but it could also manipulate and make it point to a different address of our code.
-  Jumping to such a capability would cause an the program to execute in a way we did not intend for it.
-  Further in order to be able to retrieve our capabilities from before transferring control, we would have to store the capabilities somewhere accessible from the return capability.
-  However, the untrusted code would have access to all this through the return capability because it gives the same authority to us as the untrusted code.
-  This is why, any reasonable capability machine must have an encapsulation mechanism to allow programs to make boundaries between security domains.
-
-  In the example, we could seal the return capability to establish a boundary.
-  Specifically, it would prevent the untrusted code from changing the target of the return capability forcing them to return to the point we specified.
-  It would also prevent the untrusted code from reading our capabilities.
-  All in all, this means that we can transfer control to untrusted code without giving up our capabilities or handing them over.
-
-  \trgcm{} does not have a direct unsealing instruction, but it is still possible to emulate a limited unsealing mechanism.
-  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
-  Now assume part of the code would like to have access to $w$ in $\rdata$.
-  Take a capability for this code, possibly by adjusting the pc, and seal it with $\sigma$.
-  Now you have the data part and the code part of a sealed capability pair, so you can $\txjmp{}{}$ to it which unseals the data capability and puts it in $\rdata$ for your code to use.
-  Sealed executable capabilities cannot be unsealed in the same way because $\txjmp{}{}$ fails if the data capability is executable.
-  However, given a sealed executable capability $\sealed{\sigma,c}$ and a set of seals that contains $\sigma$, we can still construct the data part of the sealed capability pair.
-  This means that we can execute the code the capability points to together with data that it was not intended to be used with.
-
-  Sealing is meant for encapsulation, but it relies on seals being kept private as illustrated by the unsealing emulation.
-  For this reason, it is important that the system is initialised such that each component has access to unique seals.
+  For now, let us pretend that there does not exist a stack and only a single invocation of untrusted code happens in the entire execution.
+
+  If we did not have an encapsulation mechanism, our only option would be to give the adversary an executable capability for the address we want them to return to.
+  The untrusted code could use the return capability as intended, but it could also manipulate it to point elsewhere in our code.
+  Jumping to such a capability could cause the trusted program to execute in an unintended and potentially insecure way.
+
+  Additionally, when the untrusted code returns,  we need to regain access to the capabilities that represent our authority, which we had access to before passing control.
+  To achieve this, we have to store them in a location that we can access after the return.
+  However, without an encapsulation mechanism, the untrusted code would have access to all this through the return capability because it gives the same authority to the untrusted code as to us after the return.
+  This is why, any reasonable capability machine must have an encapsulation mechanism to allow programs to set up boundaries between security domains.
+
+  In \trgcm{}, we can seal the return capability to establish such a boundary.
+  Specifically, we can seal the return capability (which points to the instruction to be executed after the return) as a pair, together with a pointer to the stack frame where we have stashed away our capabilities (see further).
+  By doing this, the untrusted code is prevented from changing the target of the return capability forcing them to return to the point we specified.
+  Additionally, they do not get access to the capabilities we stashed away.
+  Nevertheless, the automatic unsealing of sealed capability pairs upon invocation will ensure that the return code can still proceed as before.
+  In other words, the sealing mechanism in \trgcm{} allows us to transfer control to untrusted code without giving up our capabilities or handing them over, and also control the locations in our code they can jump back to.
+
+  It is worth pointing out that \trgcm{} does not have a direct unsealing instruction to extract a capability from its sealed version when the seal is available, but one can be emulated.
+  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals capability $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
+  It is possible to extract $w$ in $\rdata$ using $\txjmp{}{}$ in the following way.
+  The idea is to use the seal to construct a sealed version of the pc capability incremented by one, and then perform an $\txjmp{}{}$ to it in combination with the sealed capability.
+  This does not work for sealed executable capabilities because $\txjmp{}{}$ fails if the data capability is executable.
+  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.
+
+  Sealing is meant for encapsulation, but it relies on seals being kept private as should be clear from the explanations so far.
+  For this reason, it is important that trusted code does not leak its seals to adversaries and that the system is initialised such that each component has access to unique seals.
   We return to this in \sectionname~\ref{sec:well-form-reas}.
@@
 We also assume a function $\encInstr{}$ to make it easy to specify programs in terms of instructions.
-Rather than specifying a decode function and an encode function, we assume that they are given with certain properties.
+Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
 The $\decInstr{} : \Word \fun \Instr$ should be surjective and injective for all non-$\fail$ instructions.
-Further, it should decode all capabilities as the $\fail$ instruction, i.e.\
+Further, it should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
 \[
@@
 \]
-The $\encInstr{} : \Instr \fun \ints$ function should be injective, and it should be defined so $\decInstr{}$ is its left inverse, i.e.
+The $\encInstr{} : \Instr \fun \ints$ function should be injective, and  $\decInstr{}$ is its left inverse, i.e.
 \[
@@
 \]
-
-These assumptions are sufficient to construct program examples and run them on the machine.
-The $\encInstr{}$ function allows us to specify the program ``abstractly'' in terms of instructions rather than machine words.
-As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
+These assumptions are sufficient to construct program examples in terms of instructions rather than machine words (using $\encInstr{}$), and run them on the machine (using the fact that $\decInstr{}$ is the left inverse of $\encInstr{}$).
+% The $\encInstr{}$ function allows us to specify the program ``abstractly''.
+% As the $\decInstr{}$ function is the left inverse of $\encInstr{}$, we can even execute the program in the operational semantics without ever worrying about what the actual encoding is.
 %This is not to say that encoding is an easy problem. There are many things to get.
@@
 For linearity we assume similar functions.
-
 Finally in the interpretation of $\tisptr{}{}$, the machine uses an encode function for word types $\encType{}$.
@@
 % Structure
-A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals\footnote{We will return to the seals in Section~\ref{sec:form-secur-with}.}, and a set of all the linear addresses (addresses governed by a linear capability).
+A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}.), and a set of all the linear addresses (addresses governed by a linear capability).
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
-The exports are words each associated with a symbol.
+An export associates a word with a symbol.
 %base = library
@@
 % Linking .
-Components are combined into new components by linking them together, as long as only one is an incomplete program with a main.
+Components are combined into new components by linking them together, as long as only one has a main function.
 Two components can be linked when their memories, seals, and linear addresses are disjoint.
@@
 For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
-The satisfied imports are removed from the import list in the resulting linked component and the exports are updated to be the exports of the two components.
+The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.
 
@@
 First, we replace stack pointers with stack capabilities.
-When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ it does not cover the caller's stack frame.
-Return pointers, on the other hand, are replaced by a pair of sealed return capabilities, as we already explained in Section~\ref{sec:purpose-sealing}.
+When a new stack frame is created, the caller provisions it with a stack capability, restricted to the appropriate range, i.e.\ not covering the caller's stack frame.
+Return pointers, on the other hand, are replaced by a pair of sealed return capabilities, as explained in Section~\ref{sec:purpose-sealing}.
 % KJAA: Again with the sealed pair. Again, this idea is unclear.
-They form an opaque closure that the callee can only jump to, and the caller's data becomes available to the caller's return code.
+They form an opaque closure that the callee can only jump to, and when that happens, the caller's data becomes available to the caller's return code.
 
@@
 Untrusted callees receive a stack capability and a return pair that they are supposed to use for the call.
-However, a malicious callee (which we will refer to as an adversary\footnote{See Section~\ref{sec:well-form-reas} for more details on our attacker model.}) can store the provided capabilities on the heap in order to use them later.
+However, a malicious callee (which we will refer to as an adversary\footnote{See Section~\ref{sec:well-form-reas} for more details on our attacker model.}) can store the provided capabilities away on the heap in order to use them later.
 Figure~\ref{fig:stack-ret-ptr-abuse} illustrates two examples of this.
@@
 % Informally explain how we prevent this using linear capabilities
-As the examples illustrate, this naive use of standard memory and object capabilities does not provide sufficient guarantees to enforce LSE and WBCF.
-The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect after their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
-Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver no longer has it or a copy of it as it is non-duplicable.} that can be used to prevent this from happening.
+As the examples illustrate, this naive use of memory and object capabilities does not suffice to enforce LSE and WBCF.
+The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect outside their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
+Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver cannot have kept a copy of it as it is non-duplicable.} that can be used to prevent this from happening.
 
@@
 One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
-The callee would have to provide this token on return to prove that it is allowed to return to the caller, and on return the token would be taken back by the caller to prevent it from being spent multiple times.
+The callee would have to present this token on return to prove that it is allowed to return to the caller, and on return the token would be taken back by the caller to prevent it from being spent multiple times.
 As it turns out, the stack capability for the unused part of the stack can be used as such a token in the following way:
 On return the callee has to give back the stack capability they were given on invocation.
-When the caller receives a stack capability back on return, they need to check that this token is actually spendable, i.e.\ check whether their stack frame is on top of the call stack or not.
-They do this by attempting to restore the stack capability from before the call by splicing the return token with the stack capability for the local stack frame which at this point has been unsealed again.
+When the caller receives a stack capability back on return, it checks that this token is actually spendable, i.e.\ whether its stack frame is on top of the logical call stack or not.
+This check can be done by attempting to recombine (splice) the return token with the stack capability for the local stack frame (which at this point has been unsealed again) in order to reconstruct the original stack capability.
 If the splice is successful, then the caller knows that the two capabilities are adjacent.
@@
 Non-empty stack frames come naturally in traditional C-like calling convention as they keep track of old stack pointers and old program counters on the stack, but in \stktokens{} these things are part of the return pair which means that a caller with no local data may only need an empty stack frame.
-This means that a caller using \stktokens{} needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
-There is also a more practical reason for a \stktokens{} caller to make sure their stack frame is non-empty: They need a bit of the stack capability in order to perform the splice that verifies the validity of the return token.
+In other words, a caller using \stktokens{} needs to take care that their stack frame is non-empty in order to reserve their spot in the return order.
+There is also a more practical reason for a \stktokens{} caller to make sure their stack frame is non-empty: They need a fragment of the stack capability in order to perform the splice that verifies the validity of the return token.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}{0.23\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \draw (-0.5,13) node {\footnotesize 0};
+      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \actadv{(0,4)}{(4.5,7)}{}
+
+      \sealsymb{(4.9,3)}
+      \lincapbrace{(4.6,2)}{(4.6,4)}{}
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+      \lincapbracebot{(4.6,4)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-a}
+  \end{subfigure}
+  \begin{subfigure}{0.24\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \inactadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
+      \node[opacity=0.5,rotate=-45] at (2.25,10) {\footnotesize Sent to trusted};
+
+      \sealsymb{(4.9,3)}
+      \lincapbrace{(4.6,2)}{(4.6,4)}{}
+      \lincapbrace{(4.6,4)}{(4.6,7)}{}
+
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+      \lincapbracebot{(4.6,7)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-b}
+  \end{subfigure}
+  \begin{subfigure}{0.24\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+      \actadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
+      \node[opacity=0.5,rotate=-45] at (2.25,11) {\footnotesize Returned to adv.};
+      \sealsymb{(4.9,3)}
+      \lincapbrace{(4.6,2)}{(4.6,4)}{}
+      \lincapbrace{(4.6,4)}{(4.6,7)}{}
+
+      \sealsymb{(4.9,8)}
+      \lincapbrace{(4.6,7)}{(4.6,9)}{}
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-c}
+  \end{subfigure}
+  \begin{subfigure}{0.24\linewidth}
+    \centering
+    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+      % recurrent parts
+      \stdstackstart[13]
+      \actsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
+
+      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
+      \begin{scope}
+        \clip (-.1,-.1) rectangle (4.6,13.1);
+        \draw[fill=gray!50] (0,9) rectangle (4.5,13.5);
+      \end{scope}
+
+      \lincapbrace{(4.6,2)}{(4.6,7)}{}
+
+      \sealsymb{(4.9,8)}
+      \lincapbrace{(4.6,7)}{(4.6,9)}{}
+
+      \fill[red,draw=red,opacity=.4] (-.1,6.9) rectangle (4.6,9.1);
+      \begin{scope}
+        \clip (4.8,1) rectangle (9,13);
+        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
+      \end{scope}
+
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:stack-base-abuse-d}
+  \end{subfigure}
+  \caption{Partial return token used to return out of order.}
+  \label{fig:stk-base-abuse}
+\end{figure}
 
@@
 Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
-This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a capability for reading, either directly or indirectly, a return seal from memory).
+This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a different capability that can be used for reading, either directly or indirectly, a return seal from memory).
 
@@
 However, \stktokens{} is a modular calling scheme: no restriction is put on adversarial components that we do not expect trusted components to meet.
-Specifically, we are going to assume that both trusted and adversarial components are initially syntactically well-formed (described in more detail in Section~\ref{sec:well-form-reas}) which basically just restrict adversarial components to not break machine guarantees initially (e.g.\ no aliases for linear capabilities or access to seals of other components).
-This means that any component can ensure WBCF and LSE by employing \stktokens{}.
+Specifically, we will only assume that both trusted and adversarial components are initially syntactically well-formed (described in more detail in Section~\ref{sec:well-form-reas}) which basically just ensures that their initial state does not break machine guarantees (e.g.\ no aliases for linear capabilities or access to seals of other components).
+This means that mutually distrusting components can ensure WBCF and LSE for themselves by employing \stktokens{}.
 
@@
 The former is sealed and used for the data part of the return pair.
+The latter is passed to the callee as the stack pointer.
 \item \label{item:return-code} Create code return capability on call: Seal the old program counter capability.
@@
 We do not include this check because it only needs to happen once between two calls, so that the check after a call suffices if the stack base is not changed subsequently.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{0.23\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \draw (-0.5,13) node {\footnotesize 0};
-      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \actadv{(0,4)}{(4.5,7)}{}
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-      \lincapbracebot{(4.6,4)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-a}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \inactadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
-      \node[opacity=0.5,rotate=-45] at (2.25,10) {\footnotesize Sent to trusted};
-
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{}
-
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-      \lincapbracebot{(4.6,7)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-b}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-      \actadv{(0,4)}{(4.5,7)}{\footnotesize Kept by adv.}
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
-      \node[opacity=0.5,rotate=-45] at (2.25,11) {\footnotesize Returned to adv.};
-      \sealsymb{(4.9,3)}
-      \lincapbrace{(4.6,2)}{(4.6,4)}{}
-      \lincapbrace{(4.6,4)}{(4.6,7)}{}
-
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{}
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-c}
-  \end{subfigure}
-  \begin{subfigure}{0.24\linewidth}
-    \centering
-    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
-      % recurrent parts
-      \stdstackstart[13]
-      \actsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
-
-      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted\\ \footnotesize stack frame 2}
-      \begin{scope}
-        \clip (-.1,-.1) rectangle (4.6,13.1);
-        \draw[fill=gray!50] (0,9) rectangle (4.5,13.5);
-      \end{scope}
-
-      \lincapbrace{(4.6,2)}{(4.6,7)}{}
-
-      \sealsymb{(4.9,8)}
-      \lincapbrace{(4.6,7)}{(4.6,9)}{}
-
-      \fill[red,draw=red,opacity=.4] (-.1,6.9) rectangle (4.6,9.1);
-      \begin{scope}
-        \clip (4.8,1) rectangle (9,13);
-        \lincapbracebot{(4.6,9)}{(4.6,13.5)}{}
-      \end{scope}
-
-    \end{tikzpicture}
-    \caption{}
-    \label{fig:stack-base-abuse-d}
-  \end{subfigure}
-  \caption{Partial return token used to return out of order.}
-  \label{fig:stk-base-abuse}
-\end{figure}
 
@@
 \begin{jversion}
-The new run-time values in the overlay semantics affect the definition of the encoding function $\encType{}$.
-All the new values correspond to concrete capabilities on the \trgcm{} machine which the encoding function must respect.
-For instance, the encoding of a stack pointer in the overlay semantics should be the same as the encoding of a linear capability on the \trgcm{} machine.
+The new run-time values in the overlay semantics are treated appropriately in functions like $\encType{}$.
+All the new values correspond to concrete capabilities on the \trgcm{} machine which the encoding function reflects.
+For instance, the encoding of a stack pointer in the overlay semantics is the same as the encoding of a linear memory capability on the \trgcm{} machine.
 \end{jversion}
@@
 In other words, they cannot do anything that doesn't correspond to something that a well-behaved component, respecting LSE and WBCF, can also do.
-More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics iff they are indistinguishable to arbitrary other components in the overlay semantics.
-
-Our formal results are complicated by the fact that they only hold on a sane initial configuration of the system and for well-behaved components that respect the basic rules of the calling convention.
+More formally, our full-abstraction result states that two trusted components are indistinguishable to arbitrary other components in the regular semantics if and only if they are indistinguishable to arbitrary other components in the overlay semantics.
+
+Our formal results are complicated by the fact that they only hold on a sane initial configuration of the system and for components that respect the basic rules of the calling convention.
 For example, the system should be set up such that seals used by components for constructing return pointers are not shared with other components.
@@
   We expect the linker to link the two programs together making sure that all the system invariants are respected.
-  We expect trusted components to be the product of a known compiler that uses \stktokens{} which means that the code must have a certain shape.
-  The proper linker behaviour and the syntactic expectations we have of trusted component are captured in a syntactic well-formedness judgement which we explain in this section.
-
-  The well-formedness judgement imposes a quite rigid structure on well-formed components: a component's code memory may contain only data and sets of seals.
+  We model a linker that relies on a certain fixed shape for all components.
+  The proper linker behaviour and the syntactic expectations we have of both trusted and untrusted components are captured in a syntactic well-formedness judgement which we explain in this section.
+
+  The well-formedness judgement imposes a quite rigid structure on components: a component's code memory may contain only data and set-of-seals capabilities.
   The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities (to the component's data memory, sealed with closure seals).
   This makes a clear separation between code and data memory as neither can contain capabilities for the other.
-  The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking a return seals that were indirectly accessible through the data structure.
-  Informally, a well-formed component respects Write-XOR-Execute which means that all capabilities for code memory at most has read-execute permission and all capabilities for data memory at most has read-write permission.
+  The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking return seals that were indirectly accessible through the data structure.
+  Informally, a well-formed component respects Write-XOR-Execute which means that all capabilities for code memory at most have read-execute permission and all capabilities for data memory at most have read-write permission.
   A component's exports should be sealed code or data capabilities, sealed with a closure seal.
@@
   Any component's code memory must fall entirely within or outside of $\ta$, so every component is either trusted or adversarial.
-  Some well-formedness requirements are specific for adversarial and trusted components.
+  Some well-formedness requirements are specific for adversarial and trusted components respectively.
   Well-formed adversarial components cannot have return seals.
-  On the other hand, well-formed trusted components can have return seals, but must make return seals available to all its calls in a way that is consistent with \stktokens{}.
-  This means that each return seal is only used for one call.
+  On the other hand, well-formed trusted components can have return seals, and must allocate unique return seals to every call site.
 \end{jversion}
@@
 \begin{jversion}
-These requirements are formally expressed by the judgement $\wdjud[\ta]{\comp}$ specifies, i.e.\ it defines when components satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
+These requirements are formally expressed by the well-formedness judgement $\wdjud[\ta]{\comp}$, i.e.\ it defines when components satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
 The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
@@
 Further, the base component should satisfy the well-formedness judgement.
-The \textsc{Base} judgement ensures sure that the component has a bit of structure.
+The \textsc{Base} rule ensures that the component has a certain structure.
 Roughly speaking, the \textsc{Base} rule requires the following:
@@
 \item \textit{Code and data memory are disjoint.}
-\item \textit{Code memory is padded with zeros ($\mspad$).}
+\item \textit{Code memory is padded with zeros ($\mspad$) and this padded memory may not be referenced.}
   This prevents code memories from being spliced together.
   If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow.
-  Further, the possible control-flows of a component suddenly depends on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
+  Further, the possible control-flows of a component suddenly depend on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
 \item \textit{All memory can either be addressed by any number of non-linear capabilities or at most one linear capability.}
@@
 \item \textit{Import and export symbols are disjoint.}
-\item \textit{One of the following two are true}
+\item \textit{One of the following is true}
   \begin{itemize}
   \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code.
-We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to seals that could be used to protect calls.}.
+We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to closure seals that could be used to protect calls.}.
   \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code.
@@
 \item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
-\item \textit{The code memory satisfies the component-code judgement (\figurename~\ref{fig:well-formed-code}).}
+\item \textit{The code memory satisfies the component-code well-formedness judgement (\figurename~\ref{fig:well-formed-code})}, allowing only seal capabilities for seals owned by the component and numbers.
+If the numbers correspond to the encoding of a call instruction, then they must reference a unique return seal in the code memory.
 \item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address.
-\item \textit{All the exports satisfy the components-export judgement (\figurename~\ref{fig:well-formed-export}).}
+It allows only numbers, linear and non-linear capabilities to memory owned by the component and values sealed with closure seals.
+\item \textit{All the exports satisfy the components-export well-formedness judgement (\figurename~\ref{fig:well-formed-export})}, requiring that they are well-formed by the component-word judgement and don't own linear memory.
 \end{itemize}
@@
 % Reasonability
-The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a component from doing something silly during execution that undermines \stktokens{}.
+The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a trusted component from doing something silly during execution that undermines \stktokens{}.
 In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
@@
 \item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
-\item It does not leak return and closure seals or means to retrieve them.
+\item It does not leak return and closure seals (or means to retrieve them).
 This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
 There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
-\item It does not store return and closure seals or means to get them.
+\item It never stores return and closure seals (or means to get them) to memory.
 By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
@@
 On the other hand, if it is not an executable capability, then this must be the data part of the pair, so it should just be a reasonable word.
-Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle. If
+Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle.
 
@@
 For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
-The adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
+However, the adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
 Can adversarial code call into trusted components?
@@
 Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
-This means that any property the source language has must somehow be there after compilation whether or not it is a property of the target language.
-If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that it behaviorally matches exactly.
-In our case, LSE and WBCF was built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
+This means that any property the source language has must somehow be there after compilation, whether or not it is a property of the target language.
+If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that exactly matches behaviorally.
+In our case, LSE and WBCF are built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
 In order to enforce these properties, components on \trgcm{} use \stktokens{}.
@@
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-Our results do not apply to dynamic compromise scenarios because they are intended to be used in the verification of a secure compiler where these scenarios are not relevant.
-\dominique{Perhaps worth pointing out that (I suspect that) our step-bounded reasonability and FTLR, actually does apply in dynamic compromise scenario's.
-  The idea would be to use the theorem multiple times with different step bounds and a different boundary between trusted and untrusted code.
-  More concretely, if a component A starts behaving non-reasonable after 10 steps, then first apply the FTLR with n = 9 and component A trusted.
-  Then apply the FTLR again with n > 10 and component A no longer trusted.
-  Then you get stronger well-bracketedness guarantees until n=9 and weaker ones after that.
-  This is a step-bounded version of the idea explained in \cite{patrignani_modular_2016}.}
+
+Another interesting relation between our work and that of \citet{Abate:2018:GCG:3243734.3243745}, is that they consider dynamic compromise scenarios, where some components start out as trusted until they are compromised and are treated as adversarial after that.
+Our full abstraction result does not apply to those scenario's because it is intended to be used in the verification of a secure compiler where such scenarios are not relevant.
+Nevertheless, we believe our Fundamental Theorem of Logical Relations (see Section~\ref{thm:ftlr}) does apply in such scenario's because it is step-bounded.
+To deal with the evolving classification of components as trusted or adversarial, we can simply apply the theorem several times with different choices of $\ta$ and taking $n$ small enough that the compromise of $\ta$ has not yet happened until that point.
+Because the theorem only requires reasonability up to $n$ steps, we get well-bracketedness guarantees up to the right step in the execution.
+This can be seen as a step-bounded version of the idea explained in \cite{patrignani_modular_2016}.
+It is an interesting observation that modular secure compilation, combined with step-bounded reasoning is sufficient to cover dynamic compromise scenario's without the need to build in dynamic compromise scenario's into the definition of secure compilation like \citet{Abate:2018:GCG:3243734.3243745}.
 

### 2020-03-11 popl-2019-journal-v/proving-full-abstraction.tex
Full pass over paper, ready to submit?

@@
 % \end{itemize}
-To prove Theorem~\ref{thm:full-abstraction}, we will essentially show that trusted components in \srccm{} are related in a certain way to their embeddings in \trgcm{}, and that untrusted \trgcm{} components are similarly related to their embeddings in \srccm{}.
+To prove Theorem~\ref{thm:full-abstraction}, we essentially show that trusted components in \srccm{} are related in a certain way to their embeddings in \trgcm{}, and that untrusted \trgcm{} components are similarly related to their embeddings in \srccm{}.
 We will then prove that these relations imply that the combined programs have the same observable behavior, i.e.\ one terminates if and only if the other does.
@@
 \begin{jversion}
-  In the previous version, we were forced to omit many aspects of the logical relation, and especially the way it deals with the linking model.
-  This was unfortunate as substantial effort goes into this work, and there are insights that may be useful for similar work.
-  Therefore the current version takes a bit longer to explain those aspects.
-  Readers who are looking for a shorter summary with less detail may refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
+  The goal of this section is not to provide full technical detail or list tedious proofs.
+  However, we believe there are many interesting aspects about our proof that we were forced to omit from the conference version, which would be worthwhile explaining to other researchers.
+  This includes, for example, our use of cross-language logical relations, the techniques we use for reasoning about seals and linear capabilities, or for the linking model.
+  In the current version, we provide a more detailed explanation of the most important of these aspects, taking care to gradually introduce the different techniques we use and to not bother the reader with tedious details.
+  Of course, this material is targeted at readers with an interest in these proof techniques and may be safely skipped by others.
+  % Dominique: I'm worried reviewers will be bothered by the fact that we recommend some readers to read the conference version, so I commented it out.
+We can still readd it in next versions if reviewers comment about this.
+  % Readers who are looking for a shorter summary with less detail may refer to the conference version of this paper for a presentation~\citep{skorstengaard_stktokens_2019} with emphasis on the intuition and fewer details.
 \end{jversion}
@@
 The possible contents of the memory also influences what the world looks like.
-For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds.
+For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds using a form of memory ownership assumptions.
 In order to relate \trgcm{} and \srccm{}, we model all the features of \srccm{} in the world which means we have to model:
@@
 In Kripke logical relations, this kind of assumption changes, that do not invalidate the relatedness of values, are modelled by the future world relation $\future$.
-The future world relation can also be thought of as the model of changes in memory over time.
-
-The memory invariants in the world should be defined, so they are monotone with respect to the future world relation.
-This means that the system assumptions capabilities rely on for safety are defined such that capabilities remain safe during execution.
-Specifically, we require the world-indexed memory invariants to be monotone in the world.
-This means that a pair of memory segments that are related now will stay related in future worlds.
+The future world relation can also be thought of as the model of allowed changes in memory over time.
+
+Safety of capabilities should be defined, so it is monotone with respect to the future world relation.
+This means that the system assumptions that capabilities rely on for safety are defined such that capabilities remain safe during execution.
+The same holds for the memory invariants stored in the world: capabilities stored in memory should remain valid under allowed changes in system assumptions.
+This means we require the world-indexed memory invariants to be monotone in the world.
+In other words, a pair of memory segments that are related now will stay related in future worlds.
 This changes the spatial regions as follows:
@@
 
-Kripke future world relations must be defined such that future worlds are extensions of the past world.
-In other words, future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
-However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on assumption anymore.
-Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory.
-We mark dropped regions as revoked in the world following \citet{ahmed_2004} and \citet{thamsborg_2011} which for all intends and purposes corresponds to actually dropping the region.
+Kripke future world relations usually allow extending worlds with extra assumptions, or take steps in protocols that the system was designed to support from the start.
+However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on this assumption any more.
+Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory and drop or replace the previous assumption.
+We mark dropped regions as revoked in the world following \citet{ahmed_2004} and \citet{thamsborg_2011} which for all intents and purposes corresponds to actually dropping the region.
 
@@
 
-Recursive domain equations are so common and the solution so standard that frameworks implement them lifting the technical burden of constructing the world by hand.
-The technique we use to solve our recursive domain equation is essentially the same used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
+Recursive domain equations are so common and the solution so standard that frameworks implement them to lift the technical burden of constructing the world by hand.
+The technique we use to solve our recursive domain equation is essentially the same as what is used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
 We believe the worlds presented here could be developed in Iris.
@@
 % Cofe's
-In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
+In this section, we explain how the world is constructed in detail (readers willing to trust us can safely skip to Section~\ref{subsubsec:joining-worlds}).
+To explain this, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
 We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for our particular world.

### 2020-03-11 popl-2019-journal-v/paper.tex
Lars's ORCID
Update the acks section

@@
 \author{Lars Birkedal}
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\orcid{0000-0003-1320-0098}          %% \orcid is optional
 \affiliation{
@@
 \begin{acks}
-  This research was supported in part by the ModuRes Sapere Aude
-  Advanced Grant from The Danish Council for Independent Research for
-  the Natural Sciences (FNU) and by Cost Action CA15123 EUTypes.
-  Dominique Devriese held a Postdoctoral Fellowship from the Research Foundation - Flanders (FWO) during most of this research.
+  This research was supported in part by the ModuRes Sapere Aude Advanced Grant
+  from The Danish Council for Independent Research for the Natural Sciences (FNU).
+  Support for an STSM was received from COST Action EUTypes (CA15123).
+  Dominique Devriese held a Postdoctoral fellowship from the Research Foundation
+  Flanders (FWO) during most of this research.
+  This research was supported in part by the Research Foundation Flanders (FWO).
 \end{acks}

### 2020-03-11 popl-2019-journal-v/paper.tex
Move the world construction to its own section.

@@
 
+\section{World Construction: solving the recursive domain equation}
+\label{subsubsec:rec-dom-eq}
+In the previous sections, we sketched what our worlds should be.
+However, the worlds we want constitute a self-referential domain equation for which no solution exists in set and domain theory.
+% Move to a different domain
+Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
+Solutions to recursive domain equations can be found using standard techniques~\citep{scott_1976,america_1989,Birkedal:2011:SKM:1926385.1926401}.
+Essentially, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
+A c.o.f.e.\ can be thought of as a set with added structure.
+Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
+
+The construction of our worlds deviates a bit from the standard recipe, because we need two future world relations and the world contains regions with one of two forms of monotonicity.
+In this section, we explain our solution of the world equation.
+To make this paper self-contained, we include a recap of relevant concepts and theorems from the literature on c.o.f.e.'s.
+This section does not provide further intuition about the worlds and can be safely skipped.
+
+Recursive domain equations are so common and the solution so standard that frameworks implement them to lift the technical burden of constructing the world by hand.
+The technique we use to solve our recursive domain equation is essentially the same as what is used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
+We believe the worlds presented here could be developed in Iris.
+Had we done the development in Iris Coq, then we would have had the world tool support of Iris and the rigorousness of Coq which, in hindsight, would have been very beneficial.
+
+% Cofe's
+In this section, we explain how the world is constructed in detail (readers willing to trust us can safely skip to Section~\ref{subsubsec:joining-worlds}).
+To explain this, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
+We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for our particular world.
+\begin{definition}[Ordered family of equivalences (o.f.e.)]
+An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and $\nequal$ is an equivalence relation for each $n$.
+The pair must satisfy the following properties:
+\begin{enumerate}
+\item $\nequal[0]$ is the total relation.
+\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
+\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
+\end{enumerate}
+\end{definition}
+\noindent An o.f.e.\ can be thought of as a set with extra structure.
+Intuitively, the index on the family of equivalences is a measure of precision.
+The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.
+On the other hand as the index decreases, the equivalences become increasingly imprecise and unable to distinguish certain elements.
+At index 0, all precision is lost and the equivalence cannot distinguish anything.
+\dominique{Introduce phrases like ``up to n steps'' etc.}
+\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
+  \label{def:cauchy-sequence}
+  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
+  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
+  in $X$.
+
+  Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
+  elements in $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
+  \begin{align*}
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
+  \end{align*}
+  or in words, the elements of the chain get arbitrarily close.
+
+  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
+  \begin{align*}
+    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
+  \end{align*}
+\end{definition}
+In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
+Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
+This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
+Functions may not only retain equivalences but make it more precise.
+Such a function is said to be \emph{contractive}.
+\begin{definition}
+  \label{def:nonexpansive-contractive-ofe}
+  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
+  The function $f$ is
+  \begin{description}
+  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
+\[
+  x \nequal_X x' \implies f(x) \nequal_Y f(x')
+\]
+  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
+\[
+  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
+  \qedhere
+\]
+  \end{description}
+\end{definition}
+\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
+There are a number of standard c.o.f.e.\ constructions for countably sets, products, sums, and functions.
+A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
+The other standard constructions are defined naturally and compose new c.o.f.e.'s from existing c.o.f.e.'s.
+
+In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
+This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
+Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
+To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
+\begin{lemma}[$\blater$ c.o.f.e.]
+  \label{def:later-cofe}
+  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
+  \[
+    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
+  \]
+  where
+  \[
+    x \nequal[n]_{\blater} x' \text{ iff } \left\{
+      \begin{array}{l}
+        n = 0 \text{ or }\\
+        x \nequal[n-1]_X x'
+      \end{array}
+    \right. \qedhere
+  \]
+\end{lemma}
+The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
+Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
+If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.
+
+Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
+The worlds we sketched have additional structure in terms of a future world relation.
+The future world relation is captured by adding a preorder to c.o.f.e.'s:
+\begin{definition}[Preordered c.o.f.e.]
+  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
+  \begin{itemize}
+  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
+  \end{itemize}
+\end{definition}
+The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
+There are also natural constructions for preordered c.o.f.e.'s that rely on the underlying preorder.
+We will need the standard construction for later, product, and sum (defined in \appendixname~\ref{app:cofe}).
+
+Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
+
+In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
+First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
+As a reminder, we sketched the memory invariants to as follows
+\[
+  \World \monfun \Rel{\MemSeg \times \MemSeg}
+\]
+To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
+We do not need a construction for the world as it will be given as the solution to the recursive equation.
+Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
+This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
+%% URel
+\begin{definition}[Uniform relation]
+  \label{def:uniform-rel-cofe}
+  Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
+  \[
+    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
+  \]
+  The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
+  \[
+    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
+  \]
+  where erasure is defined as
+  \[
+    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
+  \]
+  to get a preordered c.o.f.e.\ we use subset as the preorder.
+\end{definition}
+The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
+This means that everything is equal at step $0$ as everything is left out.
+
+Next, we define a c.o.f.e.\ of monotone, non-expansive functions
+\begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
+\label{def:monnefun-cofe}
+  Given a preordered c.o.f.e.
+$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where
+  \[
+    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
+  \]
+\end{definition}
+In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\  and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
+
+Next, we construct the seal invariant.
+As a reminder, we sketched the seal invariant as
+\[
+  \Seal \parfun \World \monnefun \URel{\SealableCaps\times\SealableCaps}
+\]
+In order to define the c.o.f.e. for this, we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
+This leaves us with the task of lifting the set of seals to a c.o.f.e.\ and defining the c.o.f.e.\ partial functions.
+Generally, all countable sets can be lifted to a c.o.f.e.\ with a family of equivalences that has full precision for all indices except index 0.
+\begin{definition}[Sets to c.o.f.e.]
+  \label{def:lift-set-cofe}
+  A countable set $S$ can be lifted to a c.o.f.e. $\cofe{S}$ where
+  \begin{itemize}
+  \item $\nequal[0]$ is the total relation
+  \item for $n > 0$ the relation $\nequal$ is the normal equality on the set
+  \end{itemize}
+\end{definition}
+This definition can be used to lift the set of all seals $\Seal$ to a c.o.f.e.
+
+The c.o.f.e.\ of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions.
+\begin{definition}[The c.o.f.e. of partial, non-expansive functions]
+\label{def:parfun-cofe}
+  Given a c.o.f.e.
+$\left(S,\left(\nequal[n]_S\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(H,\left(\nequal[n]_H\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(S \parfun H,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
+  \[
+    h \nequal h' \text{ iff } \dom(h) = \dom(h') \wedge \forall w \in \dom(h) \ldotp h(w) \nequal_H h'(w)
+  \]
+\end{definition}
+The seal invariant is defined using Definition~\ref{def:lift-set-cofe} to lift $\SealableCaps$ to a c.of.e.\, defining a c.o.f.e.\ similar to the one for memory invariants, and finally combining the two with Definition~\ref{def:parfun-cofe}.
+
+Next, we want to define the regions of the world.
+As mentioned, our regions have the added structure of the future region relation, so we need to define them as a preordered c.o.f.e.
+The future region was presented in Section~\ref{subsubsec:ft-and-revocation}.
+On shared regions, the relation is equality which makes it straight forward to define:
+\begin{definition}[Shared region preordered c.o.f.e.\ ]
+  \label{def:pure-reg-p-cofe}
+  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of shared regions as
+  \[
+  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},=\right)
+  \]
+  where the family of equivalences distributes to the underlying c.o.f.e.'s, i.e. given shared regions $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
+  \[
+  (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
+  \]
+\end{definition}
+Definition~\ref{def:pure-reg-p-cofe} gives us the shared region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
+
+Next, we define the preordered c.o.f.e.\ for spatial regions.
+The future region relation on spatial regions is non-trivial, so the definition of the preordered c.o.f.e.\ is slightly more involved.
+\begin{definition}[Spatial region preordered c.o.f.e.\ ]
+  \label{def:spatial-reg-p-cofe}
+  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
+  \[
+    \Regions = \{\spatial\}\times \var{MemInv} \cup \{\spatialo\} \times \var{MemInv'} \cup \{\revoked\}
+  \]
+  define the preordered c.o.f.e.\ of shared regions as
+  \[
+  \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
+  \]
+  where the preorder $\future$ is defined as
+  \begin{mathpar}
+  \inferrule{ r \in \Regions }{ r \future r }
+  \and
+  \inferrule{ }{ \revoked \future (\spatial,\_)}
+  \and
+  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
+  \end{mathpar}
+  and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
+  for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
+  \begin{itemize}
+  \item $n=0$
+  \item $n > 0$ and one of the following is true
+    \begin{itemize}
+    \item $r = r' = \revoked$
+    \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
+    \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
+    \end{itemize}
+  \end{itemize}
+\end{definition}
+Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:spatial-reg-p-cofe} define the preordered c.o.f.e.'s necessary for the regions in our worlds.
+
+Next, we need to define the three sub-worlds for heap, encapsulated stack frames, and free stack.
+The definitions of the sub-worlds are very similar, so we can capture the essence of all three sub worlds in one definition.
+\begin{definition}[Sub-world preordered c.o.f.e.]
+  \label{def:sub-world-p-cofe}
+  Given preordered c.o.f.e. $\left(\var{Reg},\left(\nequal[n]_\var{Reg}\right)_{n=0}^{\infty},\future_\var{Reg}\right)$ define the sub-world preordered c.o.f.e.\ as
+  \[
+    \left(\nats \parfun \var{Reg},\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+  \]
+  where the family of equivalences is defined by given $W,W' \in \nats \parfun \var{Reg}$,
+  \[
+    W \nequal W' \text{ iff } \dom(W) = \dom(W') \wedge \forall r \in \dom(W)\ldotp W(r) \nequal W'(r)
+  \]
+  and the future world relation is defined by given $W,W' \in \nats \parfun \var{Reg}$,
+  \[
+  W' \future W \text{ iff } \left\{
+    \array{l}
+    \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
+    \quad \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'.i(m(r)) \future W(r)
+     \endarray
+  \right.
+\]
+\end{definition}
+The free stack sub-world preordered c.o.f.e.\ is constructed from the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) and the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
+The result is the sub-world for the free stack
+\[
+  \Worldfs = \nats \parfun \Regions
+\]
+
+Next, we define the call stack sub-world.
+To this end, we use Definition~\ref{def:spatial-reg-p-cofe} to get the spatial region which we need to combine this with an address.
+We lift the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combine the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with the standard product construction.
+All in all, we get the a preordered c.o.f.e.\
+\[
+  \Regions \times \Addr
+\]
+which we use with Definition~\ref{def:sub-world-p-cofe} to get the preordered c.o.f.e.\ for the stack sub-world:
+\[
+  \Worlds = \nats \parfun \Regions \times \Addr
+\]
+
+Finally, we define the sub-world for the heap.
+To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) using the standard union preordered c.o.f.e.\ construction.
+The result is a preordered c.o.f.e.\
+\[
+  \Regions \cup \Regionh
+\]
+which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
+\[
+  \Worldh = \nats \parfun (\Regions \cup \Regionh)
+\]
+
+With the three sub-worlds defined, we combine them with the standard product construction for preordered c.o.f.e.'s to get our worlds.
+\[
+  \Worldh \times \Worlds \times \Worldfs
+\]
+This is the world, we would like to work with, but we still don't have a solution for the recursive domain equation.
+In fact, when we defined the regions, we just left world-index in the memory and seal invariants as an unspecified variable, which means that the above world has them as a variable.
+In order to find a solution, we still need to guard the recursive occurrence.
+To this end, we simply use the construction for later preordered c.o.f.e.'s in Definition~\ref{def:later-p-cofe} with the above to get.
+\[
+  \blater (\Worldh \times \Worlds \times \Worldfs)
+\]
+Now the recursive occurrence is guarded which means that there is a solution to the recursive domain equation.
+\begin{theorem}
+  \label{thm:rec-dom-eq-sol}
+  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
+  \[
+    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
+  \]
+  and for $\hat{W},\hat{W}' \in \Wor$
+  \[
+    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
+  \]
+  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
+\[
+  \Worldh = \RegionName \parfun (\Regions \cup \Regionh)
+\]
+and
+\[
+  \Worlds = \RegionName \parfun (\Regions \times \Addr)
+\]
+and
+\[
+  \Worldfs = \RegionName \parfun \Regions
+\]
+where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
+\begin{multline*}
+  \Regionh = 
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
+\end{multline*}
+and
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \cup\\
+    \{\revoked\}
+  \end{array} \right.
+\]
+\end{theorem}
+Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,birkedal_taste_2014} to construct the solution $\Wor$ to the recursive equation.
+Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
+This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
+This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
+\[
+  \World = \Worldh \times \Worlds \times \Worldfs
+\]
+
+%% Lau: not sure whether the following remarks should be included:
+% Note that the later placement could have been different as it just needs to guard the recursive occurrence at some level.
+% It could, for instance, have been directly on the recursive occurrence.
+% By moving the later, we would also move the place we have to go under the later in our later definitions affecting how it is working with them.
+
+% In worlds that has a simple future world relation and no future region relation, it may be an advantage to define the regions in terms of a self-referential equation, solve that, and define the worlds with the solution.
+% The advantage of this approach is that it can be done entirely in c.o.f.e.'s.
+
 \section{Discussion}

### 2020-03-11 popl-2019-journal-v/proving-full-abstraction.tex
Move the world construction to its own section.

@@
 The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.\dominique{I wonder if the renaming function obviates the need for the ``revoked'' trick.}
-
-\subsubsection{Solving the recursive domain equation}
-\label{subsubsec:rec-dom-eq}
-In the previous sections, we sketched what our worlds should be.
-However, the worlds we want constitute a self-referential domain equation for which no solution exists in set and domain theory.
-% Move to a different domain
-Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
-Solutions to recursive domain equations can be found using standard techniques~\citep{scott_1976,america_1989,Birkedal:2011:SKM:1926385.1926401}.
-Essentially, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
-A c.o.f.e.\ can be thought of as a set with added structure.
-Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
-In this section, we present some of the technical details on solving our world equation as well as the final world equation (Theorem\ref{thm:rec-dom-eq-sol}).
-This section does not provide further intuition about the worlds and can be skipped.
-
-Recursive domain equations are so common and the solution so standard that frameworks implement them to lift the technical burden of constructing the world by hand.
-The technique we use to solve our recursive domain equation is essentially the same as what is used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
-We believe the worlds presented here could be developed in Iris.
-Had we done the development in Iris Coq, then we would have had the world tool support of Iris and the rigorousness of Coq which, in hindsight, would have been very beneficial.
-
-% Cofe's
-In this section, we explain how the world is constructed in detail (readers willing to trust us can safely skip to Section~\ref{subsubsec:joining-worlds}).
-To explain this, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
-We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for our particular world.
-\begin{definition}[Ordered family of equivalences (o.f.e.)]
-An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and $\nequal$ is an equivalence relation for each $n$.
-The pair must satisfy the following properties:
-\begin{enumerate}
-\item $\nequal[0]$ is the total relation.
-\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
-\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
-\end{enumerate}
-\end{definition}
-\noindent An o.f.e.\ can be thought of as a set with extra structure.
-Intuitively, the index on the family of equivalences is a measure of precision.
-The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.
-On the other hand as the index decreases, the equivalences become increasingly imprecise and unable to distinguish certain elements.
-At index 0, all precision is lost and the equivalence cannot distinguish anything.
-\dominique{Introduce phrases like ``up to n steps'' etc.}
-\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
-  \label{def:cauchy-sequence}
-  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
-  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
-  in $X$.
-
-  Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
-  elements in $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
-  \begin{align*}
-    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
-  \end{align*}
-  or in words, the elements of the chain get arbitrarily close.
-
-  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
-  \begin{align*}
-    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
-  \end{align*}
-\end{definition}
-In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
-Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
-This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
-Functions may not only retain equivalences but make it more precise.
-Such a function is said to be \emph{contractive}.
-\begin{definition}
-  \label{def:nonexpansive-contractive-ofe}
-  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
-  The function $f$ is
-  \begin{description}
-  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
-\[
-  x \nequal_X x' \implies f(x) \nequal_Y f(x')
-\]
-  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
-\[
-  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
-  \qedhere
-\]
-  \end{description}
-\end{definition}
-\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
-There are a number of standard c.o.f.e.\ constructions for countably sets, products, sums, and functions.
-A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
-The other standard constructions are defined naturally and compose new c.o.f.e.'s from existing c.o.f.e.'s.
-
-In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
-This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
-Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
-To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
-\begin{lemma}[$\blater$ c.o.f.e.]
-  \label{def:later-cofe}
-  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
-  \[
-    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
-  \]
-  where
-  \[
-    x \nequal[n]_{\blater} x' \text{ iff } \left\{
-      \begin{array}{l}
-        n = 0 \text{ or }\\
-        x \nequal[n-1]_X x'
-      \end{array}
-    \right. \qedhere
-  \]
-\end{lemma}
-The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
-Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
-If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.
-
-Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
-The worlds we sketched have additional structure in terms of a future world relation.
-The future world relation is captured by adding a preorder to c.o.f.e.'s:
-\begin{definition}[Preordered c.o.f.e.]
-  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
-  \begin{itemize}
-  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
-  \end{itemize}
-\end{definition}
-The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
-There are also natural constructions for preordered c.o.f.e.'s that rely on the underlying preorder.
-We will need the standard construction for later, product, and sum (defined in \appendixname~\ref{app:cofe}).
-
-Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
-
-In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
-First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
-As a reminder, we sketched the memory invariants to as follows
-\[
-  \World \monfun \Rel{\MemSeg \times \MemSeg}
-\]
-To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
-We do not need a construction for the world as it will be given as the solution to the recursive equation.
-Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
-This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
-%% URel
-\begin{definition}[Uniform relation]
-  \label{def:uniform-rel-cofe}
-  Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
-  \[
-    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
-  \]
-  The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
-  \[
-    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
-  \]
-  where erasure is defined as
-  \[
-    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
-  \]
-  to get a preordered c.o.f.e.\ we use subset as the preorder.
-\end{definition}
-The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
-This means that everything is equal at step $0$ as everything is left out.
-
-Next, we define a c.o.f.e.\ of monotone, non-expansive functions
-\begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
-\label{def:monnefun-cofe}
-  Given a preordered c.o.f.e.
-$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where
-  \[
-    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
-  \]
-\end{definition}
-In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\  and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
-
-Next, we construct the seal invariant.
-As a reminder, we sketched the seal invariant as
-\[
-  \Seal \parfun \World \monnefun \URel{\SealableCaps\times\SealableCaps}
-\]
-In order to define the c.o.f.e. for this, we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
-This leaves us with the task of lifting the set of seals to a c.o.f.e.\ and defining the c.o.f.e.\ partial functions.
-Generally, all countable sets can be lifted to a c.o.f.e.\ with a family of equivalences that has full precision for all indices except index 0.
-\begin{definition}[Sets to c.o.f.e.]
-  \label{def:lift-set-cofe}
-  A countable set $S$ can be lifted to a c.o.f.e. $\cofe{S}$ where
-  \begin{itemize}
-  \item $\nequal[0]$ is the total relation
-  \item for $n > 0$ the relation $\nequal$ is the normal equality on the set
-  \end{itemize}
-\end{definition}
-This definition can be used to lift the set of all seals $\Seal$ to a c.o.f.e.
-
-The c.o.f.e.\ of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions.
-\begin{definition}[The c.o.f.e. of partial, non-expansive functions]
-\label{def:parfun-cofe}
-  Given a c.o.f.e.
-$\left(S,\left(\nequal[n]_S\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(H,\left(\nequal[n]_H\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(S \parfun H,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
-  \[
-    h \nequal h' \text{ iff } \dom(h) = \dom(h') \wedge \forall w \in \dom(h) \ldotp h(w) \nequal_H h'(w)
-  \]
-\end{definition}
-The seal invariant is defined using Definition~\ref{def:lift-set-cofe} to lift $\SealableCaps$ to a c.of.e.\, defining a c.o.f.e.\ similar to the one for memory invariants, and finally combining the two with Definition~\ref{def:parfun-cofe}.
-
-Next, we want to define the regions of the world.
-As mentioned, our regions have the added structure of the future region relation, so we need to define them as a preordered c.o.f.e.
-The future region was presented in Section~\ref{subsubsec:ft-and-revocation}.
-On shared regions, the relation is equality which makes it straight forward to define:
-\begin{definition}[Shared region preordered c.o.f.e.\ ]
-  \label{def:pure-reg-p-cofe}
-  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of shared regions as
-  \[
-  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},=\right)
-  \]
-  where the family of equivalences distributes to the underlying c.o.f.e.'s, i.e. given shared regions $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
-  \[
-  (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
-  \]
-\end{definition}
-Definition~\ref{def:pure-reg-p-cofe} gives us the shared region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
-
-Next, we define the preordered c.o.f.e.\ for spatial regions.
-The future region relation on spatial regions is non-trivial, so the definition of the preordered c.o.f.e.\ is slightly more involved.
-\begin{definition}[Spatial region preordered c.o.f.e.\ ]
-  \label{def:spatial-reg-p-cofe}
-  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
-  \[
-    \Regions = \{\spatial\}\times \var{MemInv} \cup \{\spatialo\} \times \var{MemInv'} \cup \{\revoked\}
-  \]
-  define the preordered c.o.f.e.\ of shared regions as
-  \[
-  \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
-  \]
-  where the preorder $\future$ is defined as
-  \begin{mathpar}
-  \inferrule{ r \in \Regions }{ r \future r }
-  \and
-  \inferrule{ }{ \revoked \future (\spatial,\_)}
-  \and
-  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
-  \end{mathpar}
-  and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
-  for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
-  \begin{itemize}
-  \item $n=0$
-  \item $n > 0$ and one of the following is true
-    \begin{itemize}
-    \item $r = r' = \revoked$
-    \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
-    \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
-    \end{itemize}
-  \end{itemize}
-\end{definition}
-Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:spatial-reg-p-cofe} define the preordered c.o.f.e.'s necessary for the regions in our worlds.
-
-Next, we need to define the three sub-worlds for heap, encapsulated stack frames, and free stack.
-The definitions of the sub-worlds are very similar, so we can capture the essence of all three sub worlds in one definition.
-\begin{definition}[Sub-world preordered c.o.f.e.]
-  \label{def:sub-world-p-cofe}
-  Given preordered c.o.f.e. $\left(\var{Reg},\left(\nequal[n]_\var{Reg}\right)_{n=0}^{\infty},\future_\var{Reg}\right)$ define the sub-world preordered c.o.f.e.\ as
-  \[
-    \left(\nats \parfun \var{Reg},\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
-  \]
-  where the family of equivalences is defined by given $W,W' \in \nats \parfun \var{Reg}$,
-  \[
-    W \nequal W' \text{ iff } \dom(W) = \dom(W') \wedge \forall r \in \dom(W)\ldotp W(r) \nequal W'(r)
-  \]
-  and the future world relation is defined by given $W,W' \in \nats \parfun \var{Reg}$,
-  \[
-  W' \future W \text{ iff } \left\{
-    \array{l}
-    \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
-    \quad \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'.i(m(r)) \future W(r)
-     \endarray
-  \right.
-\]
-\end{definition}
-The free stack sub-world preordered c.o.f.e.\ is constructed from the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) and the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
-The result is the sub-world for the free stack
-\[
-  \Worldfs = \nats \parfun \Regions
-\]
-
-Next, we define the call stack sub-world.
-To this end, we use Definition~\ref{def:spatial-reg-p-cofe} to get the spatial region which we need to combine this with an address.
-We lift the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combine the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with the standard product construction.
-All in all, we get the a preordered c.o.f.e.\
-\[
-  \Regions \times \Addr
-\]
-which we use with Definition~\ref{def:sub-world-p-cofe} to get the preordered c.o.f.e.\ for the stack sub-world:
-\[
-  \Worlds = \nats \parfun \Regions \times \Addr
-\]
-
-Finally, we define the sub-world for the heap.
-To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) using the standard union preordered c.o.f.e.\ construction.
-The result is a preordered c.o.f.e.\
-\[
-  \Regions \cup \Regionh
-\]
-which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
-\[
-  \Worldh = \nats \parfun (\Regions \cup \Regionh)
-\]
-
-With the three sub-worlds defined, we combine them with the standard product construction for preordered c.o.f.e.'s to get our worlds.
-\[
-  \Worldh \times \Worlds \times \Worldfs
-\]
-This is the world, we would like to work with, but we still don't have a solution for the recursive domain equation.
-In fact, when we defined the regions, we just left world-index in the memory and seal invariants as an unspecified variable, which means that the above world has them as a variable.
-In order to find a solution, we still need to guard the recursive occurrence.
-To this end, we simply use the construction for later preordered c.o.f.e.'s in Definition~\ref{def:later-p-cofe} with the above to get.
-\[
-  \blater (\Worldh \times \Worlds \times \Worldfs)
-\]
-Now the recursive occurrence is guarded which means that there is a solution to the recursive domain equation.
-\begin{theorem}
-  \label{thm:rec-dom-eq-sol}
-  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
-  \[
-    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
-  \]
-  and for $\hat{W},\hat{W}' \in \Wor$
-  \[
-    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
-  \]
-  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
-\[
-  \Worldh = \RegionName \parfun (\Regions \cup \Regionh)
-\]
-and
-\[
-  \Worlds = \RegionName \parfun (\Regions \times \Addr)
-\]
-and
-\[
-  \Worldfs = \RegionName \parfun \Regions
-\]
-where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
-\begin{multline*}
-  \Regionh = 
-  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
-  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
-\end{multline*}
-and
-\[
-  \Regions = \left\{
-  \begin{array}{l}
-    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \cup\\
-    \{\revoked\}
-  \end{array} \right.
-\]
-\end{theorem}
-Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,birkedal_taste_2014} to construct the solution $\Wor$ to the recursive equation.
-Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
-This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
-This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
-\[
-  \World = \Worldh \times \Worlds \times \Worldfs
-\]
-
-%% Lau: not sure whether the following remarks should be included:
-% Note that the later placement could have been different as it just needs to guard the recursive occurrence at some level.
-% It could, for instance, have been directly on the recursive occurrence.
-% By moving the later, we would also move the place we have to go under the later in our later definitions affecting how it is working with them.
-
-% In worlds that has a simple future world relation and no future region relation, it may be an advantage to define the regions in terms of a self-referential equation, solve that, and define the worlds with the solution.
-% The advantage of this approach is that it can be done entirely in c.o.f.e.'s.
 

### 2020-03-11 popl-2019-journal-v/paper.tex
rephrase title
Reword a phrase about differences with the conference version.

@@
   The added details include a better motivation of sealing (\sectionname~\ref{sec:purpose-sealing}), the requirements of well-formedness (\sectionname~\ref{sec:well-form-reas}) and reasonability of components (\sectionname~\ref{sec:reasonable-components}).
-  The section on proving full-abstraction (\sectionname~\ref{sec:fa-proof}) has been rewritten with details and explanations about the formal method used for the full-abstraction proof.
-  This includes a description and motivation of the Kripke world (\sectionname~\ref{subsec:worlds}) and logical relation (\sectionname~\ref{subsec:logical-relation}).
+  The section on proving full-abstraction (\sectionname~\ref{sec:fa-proof}) has been rewritten to better explain the method used for the full-abstraction proof.
+  This includes a description and motivation of the Kripke worlds (\sectionname~\ref{subsec:worlds}) and logical relation (\sectionname~\ref{subsec:logical-relation}) that we use to do this.
   This paper is accompanied by a technical report~\citep{technical_report_popl} with technical details and proofs.
@@
 
-\section{World Construction: solving the recursive domain equation}
+\section{Constructing Worlds: Solving the Recursive Domain Equation}
 \label{subsubsec:rec-dom-eq}

### 2020-03-11 popl-2019-journal-v/paper.tex
Point out in the outline that some sections are intended for experts.

@@
   This paper is accompanied by a technical report~\citep{technical_report_popl} with technical details and proofs.
+
+  Generally, we have only added material that we believe is valuable to some readers, and we have worked hard to explain the more technical material and make it digestible.
+  Additionally, while Sections~\ref{sec:cap-mach-w-seal-and-lin}, \ref{sec:stktokens-explained},~\ref{sec:form-secur-with},~\ref{sec:discussion} and~\ref{sec:related-work} are intended for all readers, we have kept the more technical sections about the proof of full abstraction separate in Sections~\ref{sec:fa-proof} and~\ref{sec:rec-dom-eq}, so that it can be easily skipped by readers who prefer to skip those explanations.
 \end{jversion}
@@
 \section{Constructing Worlds: Solving the Recursive Domain Equation}
-\label{subsubsec:rec-dom-eq}
+\label{sec:rec-dom-eq}
 In the previous sections, we sketched what our worlds should be.
@@
 \section{Related Work}
+\label{sec:related-work}
 In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation or the linear capabilities we use to do it.

### 2020-03-23 popl-2019-journal-v/paper.tex
Remove the section on constructing the Kripke world and replace it by a reference to Lau's PhD.

@@
 
-\section{Constructing Worlds: Solving the Recursive Domain Equation}
-\label{sec:rec-dom-eq}
-In the previous sections, we sketched what our worlds should be.
-However, the worlds we want constitute a self-referential domain equation for which no solution exists in set and domain theory.
-% Move to a different domain
-Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
-Solutions to recursive domain equations can be found using standard techniques~\citep{scott_1976,america_1989,Birkedal:2011:SKM:1926385.1926401}.
-Essentially, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
-A c.o.f.e.\ can be thought of as a set with added structure.
-Specifically, c.o.f.e.'s have a step-index that must satisfy a set of conditions which provide sufficient structure for a solution to exist for the recursive domain equation.
-
-The construction of our worlds deviates a bit from the standard recipe, because we need two future world relations and the world contains regions with one of two forms of monotonicity.
-In this section, we explain our solution of the world equation.
-To make this paper self-contained, we include a recap of relevant concepts and theorems from the literature on c.o.f.e.'s.
-This section does not provide further intuition about the worlds and can be safely skipped.
-
-Recursive domain equations are so common and the solution so standard that frameworks implement them to lift the technical burden of constructing the world by hand.
-The technique we use to solve our recursive domain equation is essentially the same as what is used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
-We believe the worlds presented here could be developed in Iris.
-Had we done the development in Iris Coq, then we would have had the world tool support of Iris and the rigorousness of Coq which, in hindsight, would have been very beneficial.
-
-% Cofe's
-In this section, we explain how the world is constructed in detail (readers willing to trust us can safely skip to Section~\ref{subsubsec:joining-worlds}).
-To explain this, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
-We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for our particular world.
-\begin{definition}[Ordered family of equivalences (o.f.e.)]
-An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and $\nequal$ is an equivalence relation for each $n$.
-The pair must satisfy the following properties:
-\begin{enumerate}
-\item $\nequal[0]$ is the total relation.
-\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
-\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
-\end{enumerate}
-\end{definition}
-\noindent An o.f.e.\ can be thought of as a set with extra structure.
-Intuitively, the index on the family of equivalences is a measure of precision.
-The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.
-On the other hand as the index decreases, the equivalences become increasingly imprecise and unable to distinguish certain elements.
-At index 0, all precision is lost and the equivalence cannot distinguish anything.
-\dominique{Introduce phrases like ``up to n steps'' etc.}
-\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
-  \label{def:cauchy-sequence}
-  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
-  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
-  in $X$.
-
-  Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
-  elements in $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
-  \begin{align*}
-    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
-  \end{align*}
-  or in words, the elements of the chain get arbitrarily close.
-
-  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
-  \begin{align*}
-    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
-  \end{align*}
-\end{definition}
-In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
-Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
-This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
-Functions may not only retain equivalences but make it more precise.
-Such a function is said to be \emph{contractive}.
-\begin{definition}
-  \label{def:nonexpansive-contractive-ofe}
-  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
-  The function $f$ is
-  \begin{description}
-  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
-\[
-  x \nequal_X x' \implies f(x) \nequal_Y f(x')
-\]
-  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
-\[
-  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
-  \qedhere
-\]
-  \end{description}
-\end{definition}
-\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
-There are a number of standard c.o.f.e.\ constructions for countably sets, products, sums, and functions.
-A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
-The other standard constructions are defined naturally and compose new c.o.f.e.'s from existing c.o.f.e.'s.
-
-In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
-This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
-Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
-To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
-\begin{lemma}[$\blater$ c.o.f.e.]
-  \label{def:later-cofe}
-  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
-  \[
-    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
-  \]
-  where
-  \[
-    x \nequal[n]_{\blater} x' \text{ iff } \left\{
-      \begin{array}{l}
-        n = 0 \text{ or }\\
-        x \nequal[n-1]_X x'
-      \end{array}
-    \right. \qedhere
-  \]
-\end{lemma}
-The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
-Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
-If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.
-
-Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
-The worlds we sketched have additional structure in terms of a future world relation.
-The future world relation is captured by adding a preorder to c.o.f.e.'s:
-\begin{definition}[Preordered c.o.f.e.]
-  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
-  \begin{itemize}
-  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
-  \end{itemize}
-\end{definition}
-The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
-There are also natural constructions for preordered c.o.f.e.'s that rely on the underlying preorder.
-We will need the standard construction for later, product, and sum (defined in \appendixname~\ref{app:cofe}).
-
-Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
-
-In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
-First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
-As a reminder, we sketched the memory invariants to as follows
-\[
-  \World \monfun \Rel{\MemSeg \times \MemSeg}
-\]
-To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
-We do not need a construction for the world as it will be given as the solution to the recursive equation.
-Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
-This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
-%% URel
-\begin{definition}[Uniform relation]
-  \label{def:uniform-rel-cofe}
-  Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
-  \[
-    \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
-  \]
-  The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
-  \[
-    A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
-  \]
-  where erasure is defined as
-  \[
-    \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
-  \]
-  to get a preordered c.o.f.e.\ we use subset as the preorder.
-\end{definition}
-The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
-This means that everything is equal at step $0$ as everything is left out.
-
-Next, we define a c.o.f.e.\ of monotone, non-expansive functions
-\begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
-\label{def:monnefun-cofe}
-  Given a preordered c.o.f.e.
-$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where
-  \[
-    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
-  \]
-\end{definition}
-In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\  and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
-
-Next, we construct the seal invariant.
-As a reminder, we sketched the seal invariant as
-\[
-  \Seal \parfun \World \monnefun \URel{\SealableCaps\times\SealableCaps}
-\]
-In order to define the c.o.f.e. for this, we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
-This leaves us with the task of lifting the set of seals to a c.o.f.e.\ and defining the c.o.f.e.\ partial functions.
-Generally, all countable sets can be lifted to a c.o.f.e.\ with a family of equivalences that has full precision for all indices except index 0.
-\begin{definition}[Sets to c.o.f.e.]
-  \label{def:lift-set-cofe}
-  A countable set $S$ can be lifted to a c.o.f.e. $\cofe{S}$ where
-  \begin{itemize}
-  \item $\nequal[0]$ is the total relation
-  \item for $n > 0$ the relation $\nequal$ is the normal equality on the set
-  \end{itemize}
-\end{definition}
-This definition can be used to lift the set of all seals $\Seal$ to a c.o.f.e.
-
-The c.o.f.e.\ of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions.
-\begin{definition}[The c.o.f.e. of partial, non-expansive functions]
-\label{def:parfun-cofe}
-  Given a c.o.f.e.
-$\left(S,\left(\nequal[n]_S\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(H,\left(\nequal[n]_H\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(S \parfun H,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
-  \[
-    h \nequal h' \text{ iff } \dom(h) = \dom(h') \wedge \forall w \in \dom(h) \ldotp h(w) \nequal_H h'(w)
-  \]
-\end{definition}
-The seal invariant is defined using Definition~\ref{def:lift-set-cofe} to lift $\SealableCaps$ to a c.of.e.\, defining a c.o.f.e.\ similar to the one for memory invariants, and finally combining the two with Definition~\ref{def:parfun-cofe}.
-
-Next, we want to define the regions of the world.
-As mentioned, our regions have the added structure of the future region relation, so we need to define them as a preordered c.o.f.e.
-The future region was presented in Section~\ref{subsubsec:ft-and-revocation}.
-On shared regions, the relation is equality which makes it straight forward to define:
-\begin{definition}[Shared region preordered c.o.f.e.\ ]
-  \label{def:pure-reg-p-cofe}
-  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of shared regions as
-  \[
-  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},=\right)
-  \]
-  where the family of equivalences distributes to the underlying c.o.f.e.'s, i.e. given shared regions $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
-  \[
-  (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
-  \]
-\end{definition}
-Definition~\ref{def:pure-reg-p-cofe} gives us the shared region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
-
-Next, we define the preordered c.o.f.e.\ for spatial regions.
-The future region relation on spatial regions is non-trivial, so the definition of the preordered c.o.f.e.\ is slightly more involved.
-\begin{definition}[Spatial region preordered c.o.f.e.\ ]
-  \label{def:spatial-reg-p-cofe}
-  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
-  \[
-    \Regions = \{\spatial\}\times \var{MemInv} \cup \{\spatialo\} \times \var{MemInv'} \cup \{\revoked\}
-  \]
-  define the preordered c.o.f.e.\ of shared regions as
-  \[
-  \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
-  \]
-  where the preorder $\future$ is defined as
-  \begin{mathpar}
-  \inferrule{ r \in \Regions }{ r \future r }
-  \and
-  \inferrule{ }{ \revoked \future (\spatial,\_)}
-  \and
-  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
-  \end{mathpar}
-  and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
-  for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
-  \begin{itemize}
-  \item $n=0$
-  \item $n > 0$ and one of the following is true
-    \begin{itemize}
-    \item $r = r' = \revoked$
-    \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
-    \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
-    \end{itemize}
-  \end{itemize}
-\end{definition}
-Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:spatial-reg-p-cofe} define the preordered c.o.f.e.'s necessary for the regions in our worlds.
-
-Next, we need to define the three sub-worlds for heap, encapsulated stack frames, and free stack.
-The definitions of the sub-worlds are very similar, so we can capture the essence of all three sub worlds in one definition.
-\begin{definition}[Sub-world preordered c.o.f.e.]
-  \label{def:sub-world-p-cofe}
-  Given preordered c.o.f.e. $\left(\var{Reg},\left(\nequal[n]_\var{Reg}\right)_{n=0}^{\infty},\future_\var{Reg}\right)$ define the sub-world preordered c.o.f.e.\ as
-  \[
-    \left(\nats \parfun \var{Reg},\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
-  \]
-  where the family of equivalences is defined by given $W,W' \in \nats \parfun \var{Reg}$,
-  \[
-    W \nequal W' \text{ iff } \dom(W) = \dom(W') \wedge \forall r \in \dom(W)\ldotp W(r) \nequal W'(r)
-  \]
-  and the future world relation is defined by given $W,W' \in \nats \parfun \var{Reg}$,
-  \[
-  W' \future W \text{ iff } \left\{
-    \array{l}
-    \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
-    \quad \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'.i(m(r)) \future W(r)
-     \endarray
-  \right.
-\]
-\end{definition}
-The free stack sub-world preordered c.o.f.e.\ is constructed from the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) and the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
-The result is the sub-world for the free stack
-\[
-  \Worldfs = \nats \parfun \Regions
-\]
-
-Next, we define the call stack sub-world.
-To this end, we use Definition~\ref{def:spatial-reg-p-cofe} to get the spatial region which we need to combine this with an address.
-We lift the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combine the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with the standard product construction.
-All in all, we get the a preordered c.o.f.e.\
-\[
-  \Regions \times \Addr
-\]
-which we use with Definition~\ref{def:sub-world-p-cofe} to get the preordered c.o.f.e.\ for the stack sub-world:
-\[
-  \Worlds = \nats \parfun \Regions \times \Addr
-\]
-
-Finally, we define the sub-world for the heap.
-To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) using the standard union preordered c.o.f.e.\ construction.
-The result is a preordered c.o.f.e.\
-\[
-  \Regions \cup \Regionh
-\]
-which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
-\[
-  \Worldh = \nats \parfun (\Regions \cup \Regionh)
-\]
-
-With the three sub-worlds defined, we combine them with the standard product construction for preordered c.o.f.e.'s to get our worlds.
-\[
-  \Worldh \times \Worlds \times \Worldfs
-\]
-This is the world, we would like to work with, but we still don't have a solution for the recursive domain equation.
-In fact, when we defined the regions, we just left world-index in the memory and seal invariants as an unspecified variable, which means that the above world has them as a variable.
-In order to find a solution, we still need to guard the recursive occurrence.
-To this end, we simply use the construction for later preordered c.o.f.e.'s in Definition~\ref{def:later-p-cofe} with the above to get.
-\[
-  \blater (\Worldh \times \Worlds \times \Worldfs)
-\]
-Now the recursive occurrence is guarded which means that there is a solution to the recursive domain equation.
-\begin{theorem}
-  \label{thm:rec-dom-eq-sol}
-  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
-  \[
-    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
-  \]
-  and for $\hat{W},\hat{W}' \in \Wor$
-  \[
-    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
-  \]
-  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
-\[
-  \Worldh = \RegionName \parfun (\Regions \cup \Regionh)
-\]
-and
-\[
-  \Worlds = \RegionName \parfun (\Regions \times \Addr)
-\]
-and
-\[
-  \Worldfs = \RegionName \parfun \Regions
-\]
-where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
-\begin{multline*}
-  \Regionh = 
-  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
-  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
-\end{multline*}
-and
-\[
-  \Regions = \left\{
-  \begin{array}{l}
-    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \cup\\
-    \{\revoked\}
-  \end{array} \right.
-\]
-\end{theorem}
-Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,birkedal_taste_2014} to construct the solution $\Wor$ to the recursive equation.
-Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
-This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
-This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
-\[
-  \World = \Worldh \times \Worlds \times \Worldfs
-\]
-
-%% Lau: not sure whether the following remarks should be included:
-% Note that the later placement could have been different as it just needs to guard the recursive occurrence at some level.
-% It could, for instance, have been directly on the recursive occurrence.
-% By moving the later, we would also move the place we have to go under the later in our later definitions affecting how it is working with them.
-
-% In worlds that has a simple future world relation and no future region relation, it may be an advantage to define the regions in terms of a self-referential equation, solve that, and define the worlds with the solution.
-% The advantage of this approach is that it can be done entirely in c.o.f.e.'s.
-
 \section{Discussion}

### 2020-03-23 popl-2019-journal-v/proving-full-abstraction.tex
Remove the section on constructing the Kripke world and replace it by a reference to Lau's PhD.

@@
 
-To a large extend, the definition of $\lrhname$ is pretty standard.
+To a large extent, the definition of $\lrhname$ is pretty standard.
 $\lrhname$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
@@
 \end{jversion}
+
+\subsection{Constructing Worlds: Solving the Recursive Domain Equation}
+\label{sec:rec-dom-eq}
+In the previous sections, we sketched what our worlds should be.
+However, the worlds we want constitute a self-referential domain equation for which no solution exists in set and domain theory.
+% Move to a different domain
+Therefore, we need to move to a different domain with enough structure for a solution to exist for recursive equations.
+Solutions to recursive domain equations can be found using standard techniques~\citep{scott_1976,america_1989,Birkedal:2011:SKM:1926385.1926401}.
+Essentially, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
+A c.o.f.e.\ can be thought of as a set with added structure, specifically a step-indexed notion of equality and a limit to every Cauchy sequence (i.e.\ they are complete in a similar sense as to how the real numbers are complete but the rationals are not).
+
+Explaining the construction of the world in detail would require a recap of the basic theory of c.o.f.e.'s.
+For conciseness, we choose to not include this here, but instead refer to the PhD thesis of Skorstengaard \cite[\S 3.5]{skorstengaard_thesis_2019}, which includes a detailed explanation.
+We only include the main result, which is the following theorem, asserting the existence of a World c.o.f.e.\ satisfying the recursive equation we encountered before.
+
+% In this section, we explain our solution of the world equation.
+% To make this paper self-contained, we include a recap of relevant concepts and theorems from the literature on c.o.f.e.'s.
+% This section does not provide further intuition about the worlds and can be safely skipped.
+
+% Recursive domain equations are so common and the solution so standard that frameworks implement them to lift the technical burden of constructing the world by hand.
+% The technique we use to solve our recursive domain equation is essentially the same as what is used internally in the program logic framework Iris~\citep{iris,iris2,iris3} to construct worlds.
+% We believe the worlds presented here could be developed in Iris.
+% Had we done the development in Iris Coq, then we would have had the world tool support of Iris and the rigorousness of Coq which, in hindsight, would have been very beneficial.
+
+% % Cofe's
+% In this section, we explain how the world is constructed in detail (readers willing to trust us can safely skip to Section~\ref{subsubsec:joining-worlds}).
+% To explain this, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
+% We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for our particular world.
+% \begin{definition}[Ordered family of equivalences (o.f.e.)]
+% An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and $\nequal$ is an equivalence relation for each $n$.
+% The pair must satisfy the following properties:
+% \begin{enumerate}
+% \item $\nequal[0]$ is the total relation.
+% \item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
+% \item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
+% \end{enumerate}
+% \end{definition}
+% \noindent An o.f.e.\ can be thought of as a set with extra structure.
+% Intuitively, the index on the family of equivalences is a measure of precision.
+% The larger the index is, the more refined the equivalences are which means that they can distinguish more elements from one another.
+% On the other hand as the index decreases, the equivalences become increasingly imprecise and unable to distinguish certain elements.
+% At index 0, all precision is lost and the equivalence cannot distinguish anything.
+% \dominique{Introduce phrases like ``up to n steps'' etc.}
+% \begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
+%   \label{def:cauchy-sequence}
+%   A \emph{complete ordered family of equivalences} is an ordered family of equivalences
+%   $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
+%   in $X$.
+
+%   Let $\cofe{X}$ be an o.f.e. and $\seq{x}$ be a sequence of
+%   elements in $X$. Then $\seq{x}$ is a \emph{Cauchy sequence} if
+%   \begin{align*}
+%     \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x_j \nequal[k] x_n
+%   \end{align*}
+%   or in words, the elements of the chain get arbitrarily close.
+
+%   An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
+%   \begin{align*}
+%     \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
+%   \end{align*}
+% \end{definition}
+% In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
+% Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
+% This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
+% Functions may not only retain equivalences but make it more precise.
+% Such a function is said to be \emph{contractive}.
+% \begin{definition}
+%   \label{def:nonexpansive-contractive-ofe}
+%   Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
+%   The function $f$ is
+%   \begin{description}
+%   \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
+% \[
+%   x \nequal_X x' \implies f(x) \nequal_Y f(x')
+% \]
+%   \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
+% \[
+%   x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
+%   \qedhere
+% \]
+%   \end{description}
+% \end{definition}
+% \noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
+% There are a number of standard c.o.f.e.\ constructions for countably sets, products, sums, and functions.
+% A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
+% The other standard constructions are defined naturally and compose new c.o.f.e.'s from existing c.o.f.e.'s.
+
+% In some sense, the standard c.o.f.e.\ constructions maintain the precision of the underlying c.o.f.e.'s.
+% This, however, does not utilise the c.o.f.e.\ structure in a way that allows us to construct a solution for a recursive domain equation.
+% Intuitively, the step of the c.o.f.e.\ needs to go down before the recursive occurrence.
+% To this end, we have the \emph{later} ($\blater$) c.o.f.e.:
+% \begin{lemma}[$\blater$ c.o.f.e.]
+%   \label{def:later-cofe}
+%   Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
+%   \[
+%     \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
+%   \]
+%   where
+%   \[
+%     x \nequal[n]_{\blater} x' \text{ iff } \left\{
+%       \begin{array}{l}
+%         n = 0 \text{ or }\\
+%         x \nequal[n-1]_X x'
+%       \end{array}
+%     \right. \qedhere
+%   \]
+% \end{lemma}
+% The later c.o.f.e.\ should be placed before the self-reference to make the index decrease. 
+% Intuitively, this gives something to recurse on in order to find a solution for the domain equation.
+% If we limit ourselves to a handful of standard constructions (product, sum, arrow, later, and partial functions from the natural numbers) to construct self-referential equations and make sure to guard every occurrence of the self reference with a later, then the equation has a solution.
+
+% Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
+% The worlds we sketched have additional structure in terms of a future world relation.
+% The future world relation is captured by adding a preorder to c.o.f.e.'s:
+% \begin{definition}[Preordered c.o.f.e.]
+%   A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
+%   \begin{itemize}
+%   \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
+%   \end{itemize}
+% \end{definition}
+% The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
+% There are also natural constructions for preordered c.o.f.e.'s that rely on the underlying preorder.
+% We will need the standard construction for later, product, and sum (defined in \appendixname~\ref{app:cofe}).
+
+% Sometimes we will refer to c.o.f.e.'s and preordered c.o.f.e.'s simply by their underlying set.
+
+% In the following, we will define a number of preordered c.o.f.e.'s that can be combined to get our world equation.
+% First, we define the c.o.f.e.\ for the memory invariants (we do not need a preorder here).
+% As a reminder, we sketched the memory invariants to as follows
+% \[
+%   \World \monfun \Rel{\MemSeg \times \MemSeg}
+% \]
+% To construct this, we need to define a c.o.f.e.\ for the relation and a c.o.f.e.\ for the monotone and non-expansive functions.
+% We do not need a construction for the world as it will be given as the solution to the recursive equation.
+% Generally, we can lift a set of relations to a c.o.f.e.\ by adding a downwards closed index to it.
+% This means that if a pair is in the relation for a specific index, then it is also in the relation for all smaller indices.
+% %% URel
+% \begin{definition}[Uniform relation]
+%   \label{def:uniform-rel-cofe}
+%   Given a set $R = X \times Y$, we define $\URel{R} \subseteq \powerset{\nats \times R}$ as
+%   \[
+%     \URel{R} = \{ A \in \powerset{\nats \times R} \mid \forall n \in \nats \ldotp \forall r \in R \ldotp \npair{r} \in A \implies \forall m \leq n \ldotp \npair[m]{r} \in A\}
+%   \]
+%   The uniform relation c.o.f.e.\ $\cofe{\URel{R}}$ has $\URel{R}$ as the underlying set and the following family of equivalences:
+%   \[
+%     A \nequal B \text{ iff } \erasen{A}{n} = \erasen{B}{n}
+%   \]
+%   where erasure is defined as
+%   \[
+%     \erasen{A}{n} \defeq \{(k,a) \in A \mid k < n\}
+%   \]
+%   to get a preordered c.o.f.e.\ we use subset as the preorder.
+% \end{definition}
+% The uniform relation construction uses the canonical family of equivalences which basically ignores everything from the step up and requires the rest to be equal.
+% This means that everything is equal at step $0$ as everything is left out.
+
+% Next, we define a c.o.f.e.\ of monotone, non-expansive functions
+% \begin{definition}[The c.o.f.e. of monotone and non-expansive functions]
+% \label{def:monnefun-cofe}
+%   Given a preordered c.o.f.e.
+$\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty},\future_W\right)$ and a preordered c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty},\future_U\right)$ , define the c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}\right)$ where
+%   \[
+%     h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
+%   \]
+% \end{definition}
+% In order to construct the memory invariant, we use Definition~\ref{def:monnefun-cofe} with the world solution as the $W$ c.o.f.e.\  and a uniform relation over pairs of memory segments (Definition~\ref{def:uniform-rel-cofe}) as the $U$ c.o.f.e.
+
+% Next, we construct the seal invariant.
+% As a reminder, we sketched the seal invariant as
+% \[
+%   \Seal \parfun \World \monnefun \URel{\SealableCaps\times\SealableCaps}
+% \]
+% In order to define the c.o.f.e. for this, we can use Definitions~\ref{def:uniform-rel-cofe} and~\ref{def:monnefun-cofe}.
+% This leaves us with the task of lifting the set of seals to a c.o.f.e.\ and defining the c.o.f.e.\ partial functions.
+% Generally, all countable sets can be lifted to a c.o.f.e.\ with a family of equivalences that has full precision for all indices except index 0.
+% \begin{definition}[Sets to c.o.f.e.]
+%   \label{def:lift-set-cofe}
+%   A countable set $S$ can be lifted to a c.o.f.e. $\cofe{S}$ where
+%   \begin{itemize}
+%   \item $\nequal[0]$ is the total relation
+%   \item for $n > 0$ the relation $\nequal$ is the normal equality on the set
+%   \end{itemize}
+% \end{definition}
+% This definition can be used to lift the set of all seals $\Seal$ to a c.o.f.e.
+
+% The c.o.f.e.\ of partial, non-expansive functions is defined similarly to the c.o.f.e.\ of monotone non-expansive functions.
+% \begin{definition}[The c.o.f.e. of partial, non-expansive functions]
+% \label{def:parfun-cofe}
+%   Given a c.o.f.e.
+$\left(S,\left(\nequal[n]_S\right)_{n=0}^{\infty}\right)$ and a c.o.f.e.\ $\left(H,\left(\nequal[n]_H\right)_{n=0}^{\infty}\right)$ , define the c.o.f.e.\ $\left(S \parfun H,\left(\nequal[n]\right)_{n=0}^{\infty} \right)$ where
+%   \[
+%     h \nequal h' \text{ iff } \dom(h) = \dom(h') \wedge \forall w \in \dom(h) \ldotp h(w) \nequal_H h'(w)
+%   \]
+% \end{definition}
+% The seal invariant is defined using Definition~\ref{def:lift-set-cofe} to lift $\SealableCaps$ to a c.of.e.\, defining a c.o.f.e.\ similar to the one for memory invariants, and finally combining the two with Definition~\ref{def:parfun-cofe}.
+
+% Next, we want to define the regions of the world.
+% As mentioned, our regions have the added structure of the future region relation, so we need to define them as a preordered c.o.f.e.
+% The future region was presented in Section~\ref{subsubsec:ft-and-revocation}.
+% On shared regions, the relation is equality which makes it straight forward to define:
+% \begin{definition}[Shared region preordered c.o.f.e.\ ]
+%   \label{def:pure-reg-p-cofe}
+%   Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of shared regions as
+%   \[
+%   \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},=\right)
+%   \]
+%   where the family of equivalences distributes to the underlying c.o.f.e.'s, i.e.
+given shared regions $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
+%   \[
+%   (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
+%   \]
+% \end{definition}
+% Definition~\ref{def:pure-reg-p-cofe} gives us the shared region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.
+
+% Next, we define the preordered c.o.f.e.\ for spatial regions.
+% The future region relation on spatial regions is non-trivial, so the definition of the preordered c.o.f.e.\ is slightly more involved.
+% \begin{definition}[Spatial region preordered c.o.f.e.\ ]
+%   \label{def:spatial-reg-p-cofe}
+%   Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
+%   \[
+%     \Regions = \{\spatial\}\times \var{MemInv} \cup \{\spatialo\} \times \var{MemInv'} \cup \{\revoked\}
+%   \]
+%   define the preordered c.o.f.e.\ of shared regions as
+%   \[
+%   \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
+%   \]
+%   where the preorder $\future$ is defined as
+%   \begin{mathpar}
+%   \inferrule{ r \in \Regions }{ r \future r }
+%   \and
+%   \inferrule{ }{ \revoked \future (\spatial,\_)}
+%   \and
+%   \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
+%   \end{mathpar}
+%   and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
+%   for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
+%   \begin{itemize}
+%   \item $n=0$
+%   \item $n > 0$ and one of the following is true
+%     \begin{itemize}
+%     \item $r = r' = \revoked$
+%     \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
+%     \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
+%     \end{itemize}
+%   \end{itemize}
+% \end{definition}
+% Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:spatial-reg-p-cofe} define the preordered c.o.f.e.'s necessary for the regions in our worlds.
+
+% Next, we need to define the three sub-worlds for heap, encapsulated stack frames, and free stack.
+% The definitions of the sub-worlds are very similar, so we can capture the essence of all three sub worlds in one definition.
+% \begin{definition}[Sub-world preordered c.o.f.e.]
+%   \label{def:sub-world-p-cofe}
+%   Given preordered c.o.f.e. $\left(\var{Reg},\left(\nequal[n]_\var{Reg}\right)_{n=0}^{\infty},\future_\var{Reg}\right)$ define the sub-world preordered c.o.f.e.\ as
+%   \[
+%     \left(\nats \parfun \var{Reg},\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
+%   \]
+%   where the family of equivalences is defined by given $W,W' \in \nats \parfun \var{Reg}$,
+%   \[
+%     W \nequal W' \text{ iff } \dom(W) = \dom(W') \wedge \forall r \in \dom(W)\ldotp W(r) \nequal W'(r)
+%   \]
+%   and the future world relation is defined by given $W,W' \in \nats \parfun \var{Reg}$,
+%   \[
+%   W' \future W \text{ iff } \left\{
+%     \array{l}
+%     \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
+%     \quad \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'.i(m(r)) \future W(r)
+%      \endarray
+%   \right.
+% \]
+% \end{definition}
+% The free stack sub-world preordered c.o.f.e.\ is constructed from the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) and the general sub-world construction in Definition~\ref{def:sub-world-p-cofe}.
+% The result is the sub-world for the free stack
+% \[
+%   \Worldfs = \nats \parfun \Regions
+% \]
+
+% Next, we define the call stack sub-world.
+% To this end, we use Definition~\ref{def:spatial-reg-p-cofe} to get the spatial region which we need to combine this with an address.
+% We lift the set of addresses to a preordered c.o.f.e.\ $A$ (Definition~\ref{def:lift-set-cofe} with equality as the preorder) and combine the preordered c.o.f.e.'s $\Addr$ and $\Regions$ with the standard product construction.
+% All in all, we get the a preordered c.o.f.e.\
+% \[
+%   \Regions \times \Addr
+% \]
+% which we use with Definition~\ref{def:sub-world-p-cofe} to get the preordered c.o.f.e.\ for the stack sub-world:
+% \[
+%   \Worlds = \nats \parfun \Regions \times \Addr
+% \]
+
+% Finally, we define the sub-world for the heap.
+% To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the spatial region preordered c.o.f.e.\ (Definition~\ref{def:spatial-reg-p-cofe}) using the standard union preordered c.o.f.e.\ construction.
+% The result is a preordered c.o.f.e.\
+% \[
+%   \Regions \cup \Regionh
+% \]
+% which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
+% \[
+%   \Worldh = \nats \parfun (\Regions \cup \Regionh)
+% \]
+
+% With the three sub-worlds defined, we combine them with the standard product construction for preordered c.o.f.e.'s to get our worlds.
+% \[
+%   \Worldh \times \Worlds \times \Worldfs
+% \]
+% This is the world, we would like to work with, but we still don't have a solution for the recursive domain equation.
+% In fact, when we defined the regions, we just left world-index in the memory and seal invariants as an unspecified variable, which means that the above world has them as a variable.
+% In order to find a solution, we still need to guard the recursive occurrence.
+% To this end, we simply use the construction for later preordered c.o.f.e.'s in Definition~\ref{def:later-p-cofe} with the above to get.
+% \[
+%   \blater (\Worldh \times \Worlds \times \Worldfs)
+% \]
+% Now the recursive occurrence is guarded which means that there is a solution to the recursive domain equation.
+\begin{theorem}
+  \label{thm:rec-dom-eq-sol}
+  There exists a complete ordered family of equivalences (c.o.f.e.) $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
+  \[
+    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
+  \]
+  and for $\hat{W},\hat{W}' \in \Wor$
+  \[
+    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
+  \]
+  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
+\[
+  \Worldh = \RegionName \parfun (\Regions \cup \Regionh)
+\]
+and
+\[
+  \Worlds = \RegionName \parfun (\Regions \times \Addr)
+\]
+and
+\[
+  \Worldfs = \RegionName \parfun \Regions
+\]
+where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
+\begin{multline*}
+  \Regionh = 
+  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
+  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
+\end{multline*}
+and
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \cup\\
+    \{\revoked\}
+  \end{array} \right.
+\]
+\end{theorem}
+Theorem~\ref{thm:rec-dom-eq-sol} uses the method of~\citet{Birkedal:2011:SKM:1926385.1926401,birkedal_taste_2014} to construct the solution $\Wor$ to the recursive equation.
+Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
+This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
+This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
+\[
+  \World = \Worldh \times \Worlds \times \Worldfs
+\]
+
+%% Lau: not sure whether the following remarks should be included:
+% Note that the later placement could have been different as it just needs to guard the recursive occurrence at some level.
+% It could, for instance, have been directly on the recursive occurrence.
+% By moving the later, we would also move the place we have to go under the later in our later definitions affecting how it is working with them.
+
+% In worlds that has a simple future world relation and no future region relation, it may be an advantage to define the regions in terms of a self-referential equation, solve that, and define the worlds with the solution.
+% The advantage of this approach is that it can be done entirely in c.o.f.e.'s.
 

### 2020-07-24 popl-2019-journal-v/discussion.tex
Introduction for the discussion section.
A few final pre-submission modifications.

@@
+With the technical results established, it's worth taking a step back and discuss our work from a moreo high-level perspective.
+In the next sections, we specifically give some more thoughts on our use of full abstraction to express WBCF and LSE, and the practical usability of \stktokens{}.
+
 % \begin{itemize}
@@
   Alternatively, it might also be possible to have a piece of central trusted code that does the stack unravelling for all stack frames.
-  To do this, this the trusted code would need to receive a copy of all return seals from the linker.
+  To do this, the trusted code would need to receive a copy of all return seals from the linker.
 
   % Say something about full-abstractions proofs and feature extensions
-  Full-abstraction is a property of the whole language.
-  In other words, a full-abstraction proof must consider all features of a language to make sure that the features don't interact in a way that breaks language abstractions.
-  If we have a fully abstract compiler and add a feature to the source and target language, then the new compiler is no necessarily fully-abstract.
+  Full abstraction is a property of the whole language.
+  In other words, a full abstraction proof must consider all features of a language to make sure that the features don't interact in a way that breaks language abstractions.
+  If we have a fully abstract compiler and add a feature to the source or target language, then the new compiler is not necessarily fully abstract anymore.
   Full abstraction would have to be proven for the new compiler to make sure that the new feature does not break existing abstractions in the language.
@@
   % new paragraph: conditional full-abstraction
-  \dominique{This paragraph should perhaps also point out the link to the dynamic compromise paper?}
-  \dominique{Then again, we come back to this later in the related work, so perhaps not necessary to discuss this here?}
-  Our full-abstraction theorem, Theorem~\ref{thm:full-abstraction}, is not pure full-abstraction as it requires the components to be reasonable and well-formed.
+  Our full-abstraction theorem, Theorem~\ref{thm:full-abstraction} only applies to components that are reasonable and well-formed.
   In other words, if we were to define a compiler phase that targets \srccm{}, then we would also have to show that every program it generates is well-formed and reasonable in order to use the full-abstraction result.
   Without the reasonability constraint, \stktokens{} would have to enforce reasonability instead.
-  That is, \stktokens{} would have to dynamically ensure that no return seals or means to get return seals are passed in calls.
-  Essentially, such checks would protect the trusted code against itself which shouldn't be necessary in the first place.
-  Instead, the compiler should generate reasonable code that never exhibits the unreasonable behaviour.
+  That is, \stktokens{} would have to dynamically ensure that no return seals or means to obtain them are passed in calls.
+  Such checks would impose a performance overhead and are in principle unnecessary for correct code.
+  We have instead chosen to rely on the compiler to generate reasonable code that never exhibits the unreasonable behaviour.
   This should be done in a compiler phase where more information about the original program is available, e.g.\ the compilation phase that commits to the low-level translation.
@@
 Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team as well as our own thoughts on the matter.
-As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
+As we understand it, the main problems for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
 Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
@@
 % If such memory could be (mostly) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst case.
-The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
+The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should be non-aliased) and just use a non-atomic load in that case.
 

### 2020-07-24 popl-2019-journal-v/paper.tex
disable todo notes, journal version annotations.
Include a "conflicts of interest declaration" as required by JFP.
Update Lau's affiliation (proposal).
A few final pre-submission modifications.

@@
 \definecolor{OliveGreen}{RGB}{60,128,49}
-\newenvironment{jversion}%{}{}
-    {\color{OliveGreen}}{}
+\newenvironment{jversion}{}{}
+    % {\color{OliveGreen}}{}
 \newenvironment{jversionsug}
@@
 
-\newcommand{\ijversion}[1]%{#1}
-    {{\color{OliveGreen} #1}}
-\newcommand{\ijversionsug}[1]%{}
-    {{\color{Blue} #1}}
+\newcommand{\ijversion}[1]{#1}
+    % {{\color{OliveGreen} #1}}
+\newcommand{\ijversionsug}[1]{}
+    % {{\color{Blue} #1}}
 
@@
 \affiliation{
-  \institution{Aarhus~University}
+  \institution{Toitware}
   \country{Denmark}                    %% \country is recommended
 }
-\email{lask@cs.au.dk}
+\authornote{Research performed while the author was working for Aarhus University.}
+\email{l030689@gmail.com}
 
@@
 
-\renewcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
-\renewcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
-\renewcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+% \renewcommand\lau[1]{{\color{purple} \sf \footnotesize {LS: #1}}\\}
+% \renewcommand\dominique[1]{{\color{purple} \sf \footnotesize {DD: #1}}\\}
+% \renewcommand\lars[1]{{\color{purple} \sf \footnotesize {LB: #1}}\\}
+\renewcommand\lau[1]{}
+\renewcommand\dominique[1]{}
+\renewcommand\lars[1]{}
 
@@
 \todo[inline]{TODOS ENABLED!!!}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016, patrignani_2019}), but a real secure compiler is yet to be built.
+Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016, patrignani_2019,barthe_formal_2019}), but real secure compilers are still a rare sight.
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
   Generally, we have only added material that we believe is valuable to some readers, and we have worked hard to explain the more technical material and make it digestible.
-  Additionally, while Sections~\ref{sec:cap-mach-w-seal-and-lin}, \ref{sec:stktokens-explained},~\ref{sec:form-secur-with},~\ref{sec:discussion} and~\ref{sec:related-work} are intended for all readers, we have kept the more technical sections about the proof of full abstraction separate in Sections~\ref{sec:fa-proof} and~\ref{sec:rec-dom-eq}, so that it can be easily skipped by readers who prefer to skip those explanations.
+  Additionally, while Sections~\ref{sec:cap-mach-w-seal-and-lin}, \ref{sec:stktokens-explained},~\ref{sec:form-secur-with},~\ref{sec:discussion} and~\ref{sec:related-work} are intended for all readers, we have kept the more technical sections about the proof of full abstraction separate in Sections~\ref{sec:fa-proof} and particularly \ref{sec:rec-dom-eq}, so that it can be easily skipped by readers who prefer to do so.
 \end{jversion}
@@
 Such a property ignores temporal properties and seems hard to use for reasoning.
-There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending}.
+There is also more and more evidence that these partial security properties are not enough to prevent realistic attacks in practice~\citep{Evans:2015:CJW:2810103.2813646,Carlini2015ControlFlowBending,van_der_veen_dynamics_2017}.
 
@@
 \citet{tsampas_2019} allow (almost) normal references that can be stored in multiple places at the same time.
-This means that their approach is more like C than what \stktokens{} has to offer.
+This means that their approach is more like C than \stktokens{}.
 As explained in Section~\ref{sec:introduction}, these approaches have the downside that they require stack clearing (including unused parts) on boundary crossings.
@@
   This research was supported in part by the Research Foundation Flanders (FWO).
+  Conflicts of Interest: None.
 \end{acks}

### 2020-09-02 popl-2019-journal-v/appendix.tex
port to JFP latex template.

@@
 \label{app:instr-interpretation}
-In this section, we present the interpretation of the \trgcm{} instructions left out of \figurename~\ref{fig:target-op-sem}.
+In this section, we present the interpretation of the \trgcm{} instructions left out of Figure~\ref{fig:target-op-sem1} and~\ref{fig:target-op-sem2}.
 
@@
 \section{\srccm{} instruction interpretation}
-In this section, we present the interpretation of the \srccm{} instructions left out of \figurename~\ref{fig:target-op-sem}.
+In this section, we present the interpretation of the \srccm{} instructions left out of Figure~\ref{fig:target-op-sem1} and~\ref{fig:target-op-sem2}.
 For the instructions that only change slightly on \srccm{} compared to \trgcm{}, we include the \srccm{} specific things in blue and the things both have in common in black.

### 2020-09-02 popl-2019-journal-v/paper.tex
Update Lau's email address.
port to JFP latex template.

@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-%% For editing double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,screen]{acmart}\settopmatter{}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10002978.10002986.10002989</concept_id>
-<concept_desc>Security and privacy~Formal security models</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10003006</concept_id>
-<concept_desc>Security and privacy~Systems security</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10010124.10010138</concept_id>
-<concept_desc>Theory of computation~Program reasoning</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[300]{Security and privacy~Formal security models}
-\ccsdesc[300]{Security and privacy~Systems security}
-\ccsdesc[300]{Theory of computation~Program reasoning}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-% \acmJournal{PACMPL}
-% \acmVolume{1}
-% \acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
-% \acmArticle{1}
-% \acmYear{2018}
-% \acmMonth{1}
-% \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-% \startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-% \setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
+\documentclass{jfp}
+
 \usepackage[utf8]{inputenc}
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\bibliographystyle{JFPlike}
 
@@
 
+% prevent a compilation failure with JFP.
+\undef{\base}
 \input{../fully-abstract-comp/preamble}
@@
 
-\renewcommand{\sectionname}{Section}
+\newcommand{\sectionname}{Section}
 
@@
 
-%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
-%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
-%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
-%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
-%%%
-%%% The following is specific to POPL '19 and the paper
-%%% 'StkTokens: Enforcing Well-Bracketed Control Flow and Stack Encapsulation using Linear Capabilities'
-%%% by Lau Skorstengaard, Dominique Devriese, and Lars Birkedal.
-%%%
-\setcopyright{rightsretained}
-\copyrightyear{2019}
-\acmJournal{PACMPL}
-\acmYear{2019}
-\acmVolume{3}
-\acmNumber{POPL}
-\acmArticle{19}
-\acmMonth{1}
-\acmPrice{}
-\acmDOI{10.1145/3290332}
-
 \begin{document}
 
-%% Title information
-\title[\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots]{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
-% \titlenote{with title note}
-% \subtitle{Fully abstract overlay semantics}
-% \subtitlenote{with subtitle note}
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
+\journaltitle{JFP}
+\cpr{Cambridge University Press}
+\doival{10.1017/xxxxx}
+
+\lefttitle{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots}
+\righttitle{Journal of Functional Programming}
+
+\totalpg{\pageref{lastpage01}}
+\jnlDoiYr{2020}
+
+\title{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities}
+
+\begin{authgrp}
   \author{Lau Skorstengaard}
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
   \affiliation{
-  \institution{Toitware}
-  \country{Denmark}                    %% \country is recommended
+    Toitware, Denmark \footnote{Research performed while the author was affiliated with Aarhus University.}\\
+    (\email{lau.skorstengaard@gmail.com})
   }
-\authornote{Research performed while the author was working for Aarhus University.}
-\email{l030689@gmail.com}
-
-%% Author with two affiliations and emails.
+
   \author{Dominique Devriese}
-\orcid{0000-0002-3862-6856}             %% \orcid is optional
+  % \orcid{0000-0002-3862-6856}
   \affiliation{
-  \institution{Vrije~Universiteit~Brussel}           %% \institution is required
-  \country{Belgium}                   %% \country is recommended
+    Vrije~Universiteit~Brussel, Belgium\\
+    (\email{dominique.devriese@vub.be})
   }
-\email{dominique.devriese@vub.be}         %% \email is recommended
 
   \author{Lars Birkedal}
-\orcid{0000-0003-1320-0098}          %% \orcid is optional
+% \orcid{0000-0003-1320-0098}          %% \orcid is optional
   \affiliation{
-  \institution{Aarhus~University}
-  \country{Denmark}                    %% \country is recommended
+    Aarhus~University, Denmark\\
+    (\email{birkedal@cs.au.dk})
   }
-\email{birkedal@cs.au.dk}
+\end{authgrp}
 
@@
 
-% %% 2012 ACM Computing Classification System (CSS) concepts
-% %% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-% \begin{CCSXML}
-% <ccs2012>
-% <concept>
-% <concept_id>10011007.10011006.10011008</concept_id>
-% <concept_desc>Software and its engineering~General programming languages</concept_desc>
-% <concept_significance>500</concept_significance>
-% </concept>
-% <concept>
-% <concept_id>10003456.10003457.10003521.10003525</concept_id>
-% <concept_desc>Social and professional topics~History of programming languages</concept_desc>
-% <concept_significance>300</concept_significance>
-% </concept>
-% </ccs2012>
-% \end{CCSXML}
-
-% \ccsdesc[500]{Software and its engineering~General programming languages}
-% \ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-\keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, fully abstract overlay semantics}
-
-\maketitle
+% \keywords{fully abstract compilation, secure compilation, capability machines, linear capabilities, well-bracketed control flow, stack frame encapsulation, fully abstract overlay semantics}
+
+\maketitle[F]
 
@@
   \arraycolsep=1.4pt
-  \begin{array}{rrcl l rrcl}
-   \addrbnf,\basebnf \in & \Addr & \defeq & \nats & \phantom{mak} & \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
-    \aendbnf \in & & & \Addr \cup \{\infty \} & & \sealendbnf \in & & & \Seal \cup \{\infty \}\\
-    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm & & &\linbnf & \defbnf & \linear \mid \normal\\
-    \vsc \in &\SealableCaps&\defbnf & \multicolumn{6}{l}{((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}}\\
-    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf} & & w \in &\Word & \defeq & \ints \uplus \Caps\\ 
-    r \in& \RegName & \defbnf & \multicolumn{6}{l}{\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots} \\
-    \reg \in &\Reg & \defeq & \RegName \fun \Word & & \mem \in&\Mem & \defeq & \Addr \fun \Word\\
-\ms \in  &\MemSeg & \defeq & \Addr \parfun \Word & & \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
-    &\Conf & \defeq & \multicolumn{6}{l}{\ExecConf \cup \{\failed\} \cup \{\halted\} }
+  \begin{array}{rrcl}
+    \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
+    \aendbnf & & \in & \Addr \cup \{\infty \}\\
+    \permbnf \in& \Perm & \defbnf & \rwx \mid \rx \mid \rw \mid \ro \mid \noperm \\
+    \linbnf && \defbnf & \linear \mid \normal\\
+    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats\\
+    \sealendbnf & & \in & \Seal \cup \{\infty \}\\
+    \vsc \in &\SealableCaps&\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
+    c \in&\Caps& \defbnf &  \SealableCaps \mid \sealed{\sigma,\scbnf}\\
+    w \in &\Word & \defeq & \ints \uplus \Caps\\ 
+    r \in& \RegName & \defbnf & \pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
+    \reg \in &\Reg & \defeq & \RegName \fun \Word \\
+    \mem \in&\Mem & \defeq & \Addr \fun \Word\\
+    \ms \in  &\MemSeg & \defeq & \Addr \parfun \Word\\
+    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
+    &\Conf & \defeq & \ExecConf \cup \{\failed\} \cup \{\halted\} 
   \end{array}
@@
 
-\begin{figure}[p]
+\begin{figure}
   \centering
@@
 \]
-  \begin{tabular}{|>{$}c<{$}|>{$}p{3.7cm}<{$}|>{\raggedright\arraybackslash}p{6.7cm}|}
+\caption{An excerpt of the operational semantics of \trgcm{} (part 1/2).}
+\label{fig:target-op-sem1}
+\end{figure}
+
+\begin{figure}
+\begin{tabular}{|>{$}c<{$}|>{$}p{3cm}<{$}|>{\raggedright\arraybackslash}p{6cm}|}
     \hline
@@
     \hline
-    \move{r}{\rn}                                & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\Phi(\rn)}$ \\
-    \hline
-    \load{r_1}{r_2}                              & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_1}{w_1}\\ &\update{\mem.\aaddr}{w_\aaddr})\endarray & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$ and $w_a = \linCons{w_1}$\\
-    \hline
-    \store{r_1}{r_2}                             & \updPcAddr{}\arraycolsep=0pt\array[t]{rl}(\Phi&\updReg{r_2}{w_2}\\ & \update{\mem.\aaddr}{\Phi(r_2)})\endarray & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
-    \hline
-    \geta{r_1}{r_2}                              & \updPcAddr{\Phi\updReg{r_1}{w}} & If $\Phi(r_2) = ((\_,\_),\_,\_,\aaddr)$ or $\Phi(r_2) = \seal{\_,\_,\aaddr}$, then $w = \aaddr$ and otherwise $w = -1$\\
-    \hline
-    \cca{r}{\rn}                                 &\updPcAddr{\Phi\updReg{r}{w}} & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
+  \move{r}{\rn}                                & \arraycolsep=0pt\array[t]{rl}\updPcAddr{}(&\\
+  \Phi&\updReg{\rn}{w_2}\\ & \updReg{r}{w_1})\endarray & $\rn \in \RegName$ and $w_1 = \Phi(\rn)$ and $w_2 = \linCons{\Phi(\rn)}$ \\
+    \hline
+  \load{r_1}{r_2}                              & \arraycolsep=0pt\array[t]{rl}\updPcAddr{}(&\\
+  \Phi&\updReg{r_1}{w_1}\\ &\update{\mem.\aaddr}{w_\aaddr})\endarray & $\Phi(r_2) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\baddr \le \aaddr \le \eaddr$ and $\perm \in \{\rwx,\rw,\rx,\ro\}$ and $w_1 = \Phi.\mem(\aaddr)$ and $\isLinear{w_1} \Rightarrow \perm \in \{\rwx,\rw\}$ and $w_a = \linCons{w_1}$\\
+    \hline
+  \store{r_1}{r_2}                             & \arraycolsep=0pt\array[t]{rl}&\updPcAddr{}(\\
+  &\ \ \ \Phi\updReg{r_2}{w_2}\\ &\ \ \ \ \ \   \update{\mem.\aaddr}{\Phi(r_2)})\endarray & $\Phi(r_1) = ((\perm,\_),\baddr,\eaddr,\aaddr)$ and $\perm \in \{\rwx,\rw\}$ and $\baddr \le \aaddr \le \eaddr$ and $w_2 = \linCons{\Phi(r_2)}$\\
+    \hline
+  \geta{r_1}{r_2}                              & \arraycolsep=0pt\array[t]{rl}\updPcAddr{}&(\\
+  &\Phi\updReg{r_1}{w})\endarray & If $\Phi(r_2) = ((\_,\_),\_,\_,\aaddr)$ or $\Phi(r_2) = \seal{\_,\_,\aaddr}$, then $w = \aaddr$ and otherwise $w = -1$\\
+    \hline
+  \cca{r}{\rn}                                 & \arraycolsep=0pt\array[t]{rl}\updPcAddr{}&(\\
+  &\Phi\updReg{r}{w})\endarray & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
     \hline
@@
     \hline
-    \cseal{r_1}{r_2}                             & \updPcAddr{\Phi\updReg{r_1}{\vsc}} & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
+  \cseal{r_1}{r_2}                             & \arraycolsep=0pt\array[t]{rl}\updPcAddr{}&(\\
+  &\Phi\updReg{r_1}{\vsc})\endarray & $\Phi(r_1) \in \SealableCaps$ and $\Phi(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $\sigma_\baddr \le \sigma \le \sigma_\eaddr$ and $\vsc = \sealed{\sigma,\Phi(r_1)}$ \\
     \hline
@@
   \end{tabular}
-\caption{An excerpt of the operational semantics of \trgcm{}}
-  \label{fig:target-op-sem}
+\caption{An excerpt of the operational semantics of \trgcm{} (part 2 of 2).}
+  \label{fig:target-op-sem2}
 \end{figure}
@@
 % opsem
-The operational semantics of \trgcm{} is displayed in Figure~\ref{fig:target-op-sem}.
+The operational semantics of \trgcm{} is displayed in
+Figures~\ref{fig:target-op-sem1} and~\ref{fig:target-op-sem2}.
 The operational semantics is defined in terms of a step relation that executes the next instruction in an executable configuration $\Phi$ which results in a new executable configuration or one of the two terminated configurations.
@@
 If both conditions are satisfied, then the word pointed to by the capability is decoded to an instruction which is interpreted relative to $\Phi$.
-The interpretations of some of the instructions are displayed in Figure~\ref{fig:target-op-sem}.
+The interpretations of some of the instructions are displayed in
+Figure~\ref{fig:target-op-sem1} and~\ref{fig:target-op-sem2}.
 In order to step through a program in memory, most of the interpretations use the function $\updPcAddr{}$ which simply updates the capability in the $\pcreg$ to point to the next memory address.
@@
   }{
-    (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = \comp_0 \bowtie (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d}) = (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
+    \begin{aligned}
+      (\comp_0'',c_{\mathrm{main},c}, c_{\mathrm{main},d}) &= \comp_0 \bowtie
+      (\comp_0',c_{\mathrm{main},c}, c_{\mathrm{main},d})\\
+      &=
+      (\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d}) \bowtie \comp_0'
+    \end{aligned}
   }
@@
   \centering
-  \begin{subfigure}{0.4\linewidth}
+  \begin{subfigure}{0.45\linewidth}
     \centering
@@
   \end{subfigure}
-  \begin{subfigure}{0.18\linewidth}
+  \begin{subfigure}{0.08\linewidth}
     \phantom{testtestes}
   \end{subfigure}
-  \begin{subfigure}{0.4\linewidth}
+  \begin{subfigure}{0.45\linewidth}
     \centering
@@
 
-\begin{figure}[htb]
+\begin{figure}
   \centering
@@
   \end{mathpar}
-  \begin{tabular}{|>{$}c<{$}|p{3.7cm}|>{\raggedright\arraybackslash}p{6.6cm}|}
+  \begin{tabular}{|>{$}c<{$}|p{3.9cm}|>{\raggedright\arraybackslash}p{5.4cm}|}
     \hline
@@
     \hline
-    \arraycolsep=0pt\array{c}
-    \scall{\srcalt{\offpc,\offsigma}}{\src{r_1}}{\src{r_2}}
-    \endarray &
+    \begin{aligned}
+      \scall{\srcalt{\offpc,\offsigma}}{}{}\\
+      ~\src{r_1}~\src{r_2}
+    \end{aligned}
+    &
         $\srcalt{\var{xjmpRes}(c_1,c_2,}$
@@
     $\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
-    $\sigma =\begin{multlined}[t]
+    $\begin{multlined}[t]\sigma =
       \var{getCallSeal}(\\
@@
     $w_1,w_2 = \linCons{\Phi.\reg(r_1,r_2)}$ and
-                        $\Phi.\reg(r_1,r_2) = \sealed{\sigma', c_1}, \sealed{\sigma', c_2}$
+                        $
+    \begin{multlined}
+      \Phi.\reg(r_1,r_2) =\\ \sealed{\sigma', c_1}, \sealed{\sigma', c_2}
+    \end{multlined}$
                         }\\
@@
     \arraycolsep=1.4pt
-    \begin{array}{l c >{\raggedright\arraybackslash}p{9.5cm}}
+    \begin{array}{l c >{\raggedright\arraybackslash}p{8.5cm}}
       \arraycolsep=0pt
@@
 \begin{figure}
-  \begin{subfigure}[h]{1.0\linewidth}
+  \begin{subfigure}[h]{.95\linewidth}
     \begin{mathpar}
@@
   \inferrule*[Right=C-Instr]{
-    \mscode(a) \in \ints\\
 {    \begin{multlined}
-      ([a \cdots a + \calllen-1] \subseteq \ta \wedge \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
+      ([a \cdots a + \calllen-1] \subseteq \ta \wedge\\ \mscode([a \cdots a + \calllen-1]) = \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2})
       \Rightarrow\\ (\mscode(a+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr} \wedge \sigma_\baddr+\offsigma \in \sigrets[\mathrm{owned}])
     \end{multlined}
-}  }{
+  }\\
+  \mscode(a) \in \ints
+}{
     \sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a
@@
 
-  \begin{subfigure}[h]{1.0\linewidth}
+  \begin{subfigure}[h]{.85\linewidth}
     \begin{mathpar}
@@
 
-  \begin{subfigure}[h]{1.0\linewidth}
+  \begin{subfigure}[h]{.8\linewidth}
     \begin{mathpar}
@@
       c_{\mathrm{main},c},c_{\mathrm{main},d}  = \sealed{\sigma, c_{\mathrm{main},c}', c_{\mathrm{main},d}'} &
-      \nonExec{c_{\mathrm{main},d}'}&
-      \reg(\pcreg,\rdata) = c_{\mathrm{main},c}', c_{\mathrm{main},d}' \\
-      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} & 
-      \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} \\
-      \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0&
+      \nonExec{c_{\mathrm{main},d}'}\\
+      \reg(\pcreg,\rdata) = c_{\mathrm{main},c}', c_{\mathrm{main},d}' &
+      \src{\reg(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\eaddr_\stk}} \\
+      \trg{\reg(\rstk) = ((\rw,\linear),\baddr_\stk,\eaddr_\stk,\eaddr_\stk)} &
+      \reg(\RegName \setminus \{\pcreg,\rdata,\rstk\}) = 0\\
       \range{\ms_\stk} = \{0\}&
@@
 
-\begin{acks}
+\section*{Acknowledgements}
   This research was supported in part by the ModuRes Sapere Aude Advanced Grant
@@
   Conflicts of Interest: None.
-\end{acks}
 
 \bibliography{references}
-
-%% Appendix
-% \appendix
-% \section{Appendix}
-
-% Text of appendix \ldots
 
@@
 \input{appendix}
+
+\label{lastpage01}
 \end{document}

### 2020-09-02 popl-2019-journal-v/proving-full-abstraction.tex
A few final pre-submission modifications.
port to JFP latex template.

@@
 Luckily, we can solve circular equations if we move to a different domain.
-For now, we will ignore the problem and return to the issue in Section~\ref{subsubsec:rec-dom-eq}.
+For now, we will ignore the problem and return to the issue in Section~\ref{sec:rec-dom-eq}.
 
@@
 \begin{definition}[$\oplus$, disjoint union of ownership]
-  \[
-    W_1 \oplus W_2 = W
-  \text{ iff }
+  $W_1 \oplus W_2 = W$ iff
+  \[
   \begin{array}[t]{l}
@@
 For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrhname$ as:
-\[
-  \npair{(\overline{\sigma},\src{ms},\ms_T)} \in \lrheap(\pwheap)(W') = 
+\begin{multline*}
+    \npair{(\overline{\sigma},\src{ms},\ms_T)} \in \lrheap(\pwheap)(W') =\\
     \left\{
@@
     \right.
-\]
+\end{multline*}
 \end{definition}
@@
 \label{def:free-stack-rel}
-\[
-  \memSatFStack{\src{\ms_\stk},\ms_T}{W} \text{ iff } 
+  \begin{multline*}
+    \memSatFStack{\src{\ms_\stk},\ms_T}{W} \text{ iff } \\
     \left\{
@@
     \right.
-\]
+  \end{multline*}
 \end{definition}
@@
   \label{def:stack-rel}
-\[
-  \memSatStack{\src{\stk},\ms_T}{W} \text{ iff } 
+  \begin{multline*}
+    \memSatStack{\src{\stk},\ms_T}{W} \text{ iff } \\
     \left\{
@@
       \right.
-\]
+  \end{multline*}
 \end{definition}
@@
   For memory segments $\ms_S$, $\ms_\stk$, and $\ms_T$, stack $\stk$, and world $W$ we define memory satisfaction as
-\[
-  \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff } 
+  \begin{multline*}
+      \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff }\\
       \left\{
@@
 \newcolumntype{M}{R@{${}\in{}$}L}
-\begin{tabular}{|M|c|p{4.8cm}|}
+\begin{tabular}{|M|c|p{3.5cm}|}
   \hline
@@
 The memory invariant is defined as follows
-\[
-  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \; \hat{W} \defeq \left\{ \npair{\stpair{\ms}{\ms}} \middle| 
+\begin{multline*}
+  H^\mathrm{sta,\square}_{\stpair{\ms}{\ms}} \; \gc \; \hat{W} \defeq\\
+  \left\{
+      \npair{\stpair{\ms}{\ms}} \middle|
       \begin{array}{l}
@@
     \right\}
-\]
+\end{multline*}
 The region also requires the static memory to contain words from the untrusted value relation.
@@
     \begin{array}{l}
-    \dom(\code) = [\baddr,\eaddr] \wedge \\
-      ([\baddr - 1, \eaddr + 1] \subseteq \ta \wedge \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
+    \dom(\code) = [\baddr,\eaddr] \wedge ([\baddr - 1, \eaddr + 1] \subseteq \ta
+      \wedge\\
+      \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
       \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
@@
 %
-    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\ ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
-    \begin{array}{l}
-      [b,e] \mathrel{\#} \ta \wedge\\
+    \hspace{-2cm}\left\{ \arraycolsep=0pt\array{l} \npair{\left(\arraycolsep=0pt\array{l}
+    \src{((\perm,\lin),\baddr,\eaddr,\aaddr)},\\
+    ((\perm,\lin),\baddr,\eaddr,\aaddr) \endarray \right)} \;\biggl|\\
+    \qquad\qquad\qquad\qquad\begin{array}{l}
+      [b,e] \mathrel{\#} \ta \wedge \perm \neq \rwx \wedge \\
       \begin{array}{r l l }
@@
         % we are excluding rwx pointers.
-        \perm \neq \rwx \wedge \\
         % \perm = \rwx &\Rightarrow& \array[t]{l}\npair{(\{\rwx,\rx\},\baddr,\eaddr)} \in \execCond{\lin,W} \wedge \\
@@
       \end{array}
-    \end{array}
+    \end{array}\endarray
     \right\}\cup \\
-    \hspace{-2cm}\left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\;
-    \begin{array}{l}
+    \hspace{-2cm}\left\{ \arraycolsep=0pt\array{l} \npair{\left(\arraycolsep=0pt\array{l} \src{\stkptr{\perm,\baddr,\eaddr,\aaddr}},\\ ((\perm,\linear),\baddr,\eaddr,\aaddr) \endarray \right)} \;\biggl|\;\\
+    \qquad\qquad\qquad\qquad\begin{array}{l}
       \begin{array}{r l l}
@@
       \end{array}
-    \end{array}
+    \end{array}\endarray
     \right\} \cup \\
@@
       \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
-      \quad \isLinear{\src{\vsc_S}} \\
-      \qquad \Rightarrow \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in \sealedCond{W,H_\sigma} \wedge\\
-      \quad \nonLinear{\src{\vsc_S}} \\
-      \qquad \Rightarrow  \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in \sealedCond{\mathit{purePart}(W),H_\sigma}\\
+      \quad \isLinear{\src{\vsc_S}} \Rightarrow
+      \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in \\
+      \qquad \qquad \qquad \qquad \sealedCond{W,H_\sigma} \wedge\\
+      \quad \nonLinear{\src{\vsc_S}} \Rightarrow
+      \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in\\ 
+      \qquad\qquad\sealedCond{\mathit{purePart}(W),H_\sigma}\\
       % This case was never used (?) so it is not necessary.
@@
   \centering
-  \[
-  \lrvtrusted[\square,\gc](W) =
+    \begin{multline*}
+      \lrvtrusted[\square,\gc](W) =\\
       \begin{array}[t]{l}
@@
           \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,\gc} \wedge\\
-      \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
+          \quad \dom(\code) \subseteq \ta \wedge \\
+          \quad [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
           \quad \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss
@@
       \end{array}
-\]
+    \end{multline*}
 
@@
     \right\}
-\\  \lrexj(W) &= \left\{ \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} \middle| 
+    \\  \lrexj(W) &= \left\{ \arraycolsep=0pt\array{l}(n, (\src{v_{c,S},v_{d,S}},\\
+    \qquad v_{c,T},v_{d,T}) )\endarray \middle| 
     \begin{array}{l}
@@
 \begin{align*}
-  \lrcomp(W) &=
-  \left\{\begin{aligned}
+  \lrcomp(W) &=\\
+  &\hspace{-.7cm}\left\{\begin{aligned}
      &\npair{\src{\var{comp}},\src{\var{comp}}} \;\mid \;\\
@@
       &\qquad\quad\text{and } \msdata' = \msdata{}[\overline{a_{\mathrm{import}} \mapsto w_{\mathrm{import}}}] \\
-      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\src{\mscode\uplus \msdata'}, \mscode\uplus\msdata')} \in \lrheap(\pwheap)(W') \text{ and }\\
+      &\qquad\quad\text{then } \npair{(\sigrets\uplus\sigcloss,\src{\mscode\uplus \msdata'}, \mscode\uplus\msdata')} \\
+      &\qquad\qquad\qquad\qquad\qquad\in\lrheap(\pwheap)(W') \text{ and }\\
       &\qquad\quad\overline{\npair{(\src{w_{\mathrm{export}}},w_{\mathrm{export}})}} \in \lrv(\purePart{W'})
@@
     \right\}\\
-  &\cup \left\{
+  &\hspace{-1cm}\cup \left\{
     \begin{multlined}
@@
 
-    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\Small Contextual equivalence preservation} (0,1);
+    \draw[out=-90,in=90,double,-implies,double equal sign distance] (0,5) to node[sloped, yshift =.7em]{\small Contextual equivalence preservation} (0,1);
   \end{tikzpicture}

### 2020-12-14 popl-2019-journal-v/discussion.tex
Add file with JFP reviews in order to collect responses, fix trivial typos already.

@@
-With the technical results established, it's worth taking a step back and discuss our work from a moreo high-level perspective.
+With the technical results established, it's worth taking a step back and discuss our work from a more high-level perspective.
 In the next sections, we specifically give some more thoughts on our use of full abstraction to express WBCF and LSE, and the practical usability of \stktokens{}.
@@
   If we have native call and return instructions in the source language, then the source language would be different from the target language, and we would have to use a non-trivial back-translation.
-  Specifically, the back-translation would need to distinguish sequences of instructions that is the translation of a call from sequences of instructions that just look like a call.
+  Specifically, the back-translation would need to distinguish sequences of instructions that are the translation of a call from sequences of instructions that just look like a call.
   With overlay semantics, this is not a concern because everything that looks like a call is interpreted as a call.

### 2020-12-14 popl-2019-journal-v/paper.tex
Oops.. not a functional pearl...
Add file with JFP reviews in order to collect responses, fix trivial typos already.

@@
 \newcommand{\seq}[2][n]{\left(#2_{#1}\right)_{#1=0}^{\infty}}
-\setlength{\belowcaptionskip}{-5pt}
+% \setlength{\belowcaptionskip}{-5pt}
 
@@
 
-\maketitle[F]
+\maketitle
 
@@
 Although the terms ``well-bracketed control flow'' and ``local state encapsulation'' sound precise, it is actually far from clear what they mean, and how best to formalize them.
-Existing formulations are either partial and not suitable for reasoning~\cite{abadi_control-flow_2005} or lack evidence of generality~\cite{skorstengaard_reasoning_2017}.
+Existing formulations are either partial and not suitable for reasoning~\citep{abadi_control-flow_2005} or lack evidence of generality~\citep{skorstengaard_reasoning_2017}.
 We propose a new formulation using a technique we call {\itshape fully abstract overlay semantics}.
@@
 To express security of our calling convention, we then show that considering the same components in the original semantics does not give adversaries additional ways to interact with them. 
-More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\cite{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
+More formally, we show that mapping a component in the well-behaved semantics to the same component in the original semantics is fully abstract~\citep{abadi_protection_1999}, i.e.\ components are indistinguishable to arbitrary adversaries in the well-behaved language iff they are indistinguishable to arbitrary adversaries in the original language.
 
@@
 %The M-Machine~\citep{Dally1997Memo59} inspired capability machine in \citet{skorstengaard_reasoning_2017} uses enter-capabilities to achieve this.
-On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\cite{watson_cheri_2015,watson_fast_2016}.
+On \trgcm{} we have CHERI-like sealed capabilities to achieve this~\citep{watson_cheri_2015,watson_fast_2016}.
 % seals
@@
 Such a capability represents the authority to seal other capabilities with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In spirit of memory capabilities, a set of seals capability has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
+In the spirit of memory capabilities, a set of seals capability has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
 % On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
@@
 %%% Domains
-Words on \trgcm{} are capabilities and data (represented by integers $\ints$).
+Words on \trgcm{} are either capabilities or data (represented by integers $\ints$).
 We assume a finite set of register names $\RegName$ containing at least the registers $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$.
@@
   However, without an encapsulation mechanism, the untrusted code would have access to all this through the return capability because it gives the same authority to the untrusted code as to us after the return.
-  This is why, any reasonable capability machine must have an encapsulation mechanism to allow programs to set up boundaries between security domains.
+  This is why any reasonable capability machine must have an encapsulation mechanism to allow programs to set up boundaries between security domains.
 
@@
   Sealing is meant for encapsulation, but it relies on seals being kept private as should be clear from the explanations so far.
-  For this reason, it is important that trusted code does not leak its seals to adversaries and that the system is initialised such that each component has access to unique seals.
+  For this reason, it is important that trusted code does not leak its seals to adversaries and that the system is initialised so that each component has access to unique seals.
   We return to this in \sectionname~\ref{sec:well-form-reas}.
@@
 % Structure
-A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}.), and a set of all the linear addresses (addresses governed by a linear capability).
+A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
@@
 % write xor exec
-How a program is initialised to create an executable configuration, is discussed in Section~\ref{sec:form-secur-with}.
+How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
 % We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
@@
 \item {\itshape reusable in secure compiler chains:} for compilers using \stktokens{}, one should be able to rely on WBCF and LSE when proving correctness and security of other compiler passes and then compose such results with ours to obtain results about the full compiler.\label{def-prop:reusable}
-\item {\itshape arguably "complete"}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
+\item {\itshape arguably ``complete''}: the formalization should arguably capture the entire meaning of WBCF and LSE and should arguably be applicable to any reasonable program.
 \label{def-prop:complete}
@@
 
-The main difference between the two semantics, is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
+The main difference between the two semantics is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
 Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
@@
 The fact that \stktokens{} enforces LSE and WBCF is then formulated as a theorem about the function that maps components in the well-behaved overlay semantics to the underlying components in the regular semantics.
-The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\cite{abadi_protection_1999}.
+The theorem states that this function constitutes a fully abstract compiler, a well-known property from the field of secure compilation~\citep{abadi_protection_1999}.
 Intuitively, the theorem states that if a trusted component interacts with (potentially malicious) components in the regular semantics, then these components have no more expressive power than components which the trusted component interacts with in the well-behaved overlay semantics.
@@
 Our formal results are complicated by the fact that they only hold on a sane initial configuration of the system and for components that respect the basic rules of the calling convention.
-For example, the system should be set up such that seals used by components for constructing return pointers are not shared with other components.
+For example, the system should be set up so that seals used by components for constructing return pointers are not shared with other components.
 We envision distributing seals as a job for the linker, so this means our results depend on the linker to do this properly.
@@
 % Reasonability
-The static guarantees given by $\wdjud[\ta]{\comp}$ makes sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a trusted component from doing something silly during execution that undermines \stktokens{}.
-In order for \stktokens{} to provide guarantees for a component, we expect it to not shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
+The static guarantees given by $\wdjud[\ta]{\comp}$ make sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a trusted component from doing something silly during execution that undermines \stktokens{}.
+In order for \stktokens{} to provide guarantees for a component, we expect it not to shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
 More precisely, we expect four things of a reasonable component:
@@
   Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly or indirectly.
-  Definition~\ref{def:reasonable-pc} defines the reasonable pc's which means that if it is plugged into a configuration with a register file filled with reasonable words, then the configuration behaves reasonably.
+  Definition~\ref{def:reasonable-pc} defines the reasonable PCs which means that if it is plugged into a configuration with a register file filled with reasonable words, then the configuration behaves reasonably.
   Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
@@
   \subsubsection{Reasonable words}
-  To provide any guarantees, \stktokens{} rely on the program to no leak return seals in any way.
+  To provide any guarantees, \stktokens{} rely on the program not to leak return seals in any way.
   But what does it mean to ``leak'' a return seal?
@@
     \end{itemize}
-    We have that $\Phi$ is reasonable up to $n$ steps.
+    we have that $\Phi$ is reasonable up to $n$ steps.
   \end{definition}
@@
       If $\Phi$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$
-      for some $\src{r_1}$ and $\src{r_2}$,
+      for any $\src{r_1}$ and $\src{r_2}$,
       then all of the following hold:
@@
           ((\perm,\lin),\baddr,\eaddr,\aaddr)$
-        \item $\Phi$ does not point to $\src{\txjmp{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$
-        \item $\Phi$ does not point to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for some $\src{r_1}$ and $\src{r_2}$, $\offpc$, $\offsigma$
+        \item $\Phi$ does not point to $\src{\txjmp{r_1}{r_2}}$ for any $\src{r_1}$ and $\src{r_2}$
+        \item $\Phi$ does not point to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ for any $\src{r_1}$ and $\src{r_2}$, $\offpc$, $\offsigma$
         \item $n' = 0$ or $\Phi'$ is reasonable up to $n'-1$ steps
@@
 Item~\ref{item:dont-store-private} and Item~\ref{item:dont-leak-private} make sure that return seals are not leaked.
-Item~\ref{item:dont-store-private} says that only reasonable words can be stored to memory.
-This means that sets of return seals or execute capabilities cannot be stored to memory by a reasonable component.
+Item~\ref{item:dont-store-private} says that only reasonable words can be stored in memory.
+This means that sets of return seals or execute capabilities cannot be stored in memory by a reasonable component.
 Intuitively, a well-formed and reasonable component has its seals available in the code memory, so it can always retrieve them from the code memory.
 In other words, it is not necessary to store them in the data memory.
-Further, by making sure that seals are not stored to memory, we can allow capabilities for data memory to be passed away if there is a need for that (for instance to have a shared buffer).
+Further, by making sure that seals are not stored in memory, we can allow capabilities for data memory to be handed out if there is a need for that (for instance to have a shared buffer).
 Item~\ref{item:dont-leak-private} makes sure that seals are not leaked when transferring control to another component (i.e. on security boundary crossings).
@@
 An execution configuration may need to do other operations than calling other code, and seals should no be leaked at any point during execution.
-For this reason, Item~\ref{item:dont-leak-private} also says that if the next step is not a call and a jump, then the next execution configuration should also be reasonable.
+For this reason, Item~\ref{item:dont-leak-private} also says that if the next step is not a call or a jump, then the next execution configuration should also be reasonable.
 
@@
 This is, however, not sufficient as it does not ensure that other parts of a program don't use the return seals.
-We do not want to specify what non-call code should look like, so we just require it to not use the call seals.
+We do not want to specify what non-call code should look like, so we just require it not to use the call seals.
 This is what Item~\ref{item:return-seal-usage} ensures.
@@
 Formally, adversarial code can contain the instructions that constitute a call.
-However, for untrusted code, \srccm{} will not execute those instructions as a "native call" but execute the individual instructions separately.
+However, for untrusted code, \srccm{} will not execute those instructions as a ``native call'' but execute the individual instructions separately.
 The callee then executes in the same stack frame as the caller, so WBCF and LSE do not follow (for that call).
@@
 In terms of security goals, the technique does not enforce local stack encapsulation.
-Also, it only enforces a weak form of control flow correctness saying that jumps stay within the program's static control flow graph~\cite{Abadi2005Theory}.
+Also, it only enforces a weak form of control flow correctness saying that jumps stay within the program's static control flow graph~\citep{Abadi2005Theory}.
 Such a property ignores temporal properties and seems hard to use for reasoning.
@@
 
-CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\cite{watson_cheri_2015}.
+CheriBSD applies a similar approach with separate per-component stacks and a trusted stack manager on a capability machine~\citep{watson_cheri_2015}.
 The authors use local capabilities to prevent components from accidentally leaking their stack pointer to other components, but there is no actual capability revocation in play.
@@
 Another interesting relation between our work and that of \citet{Abate:2018:GCG:3243734.3243745}, is that they consider dynamic compromise scenarios, where some components start out as trusted until they are compromised and are treated as adversarial after that.
-Our full abstraction result does not apply to those scenario's because it is intended to be used in the verification of a secure compiler where such scenarios are not relevant.
-Nevertheless, we believe our Fundamental Theorem of Logical Relations (see Section~\ref{thm:ftlr}) does apply in such scenario's because it is step-bounded.
+Our full abstraction result does not apply to those scenarios because it is intended to be used in the verification of a secure compiler where such scenarios are not relevant.
+Nevertheless, we believe our Fundamental Theorem of Logical Relations (see Section~\ref{thm:ftlr}) does apply in such scenarios because it is step-bounded.
 To deal with the evolving classification of components as trusted or adversarial, we can simply apply the theorem several times with different choices of $\ta$ and taking $n$ small enough that the compromise of $\ta$ has not yet happened until that point.
 Because the theorem only requires reasonability up to $n$ steps, we get well-bracketedness guarantees up to the right step in the execution.
-This can be seen as a step-bounded version of the idea explained in \cite{patrignani_modular_2016}.
-It is an interesting observation that modular secure compilation, combined with step-bounded reasoning is sufficient to cover dynamic compromise scenario's without the need to build in dynamic compromise scenario's into the definition of secure compilation like \citet{Abate:2018:GCG:3243734.3243745}.
+This can be seen as a step-bounded version of the idea explained in \citep{patrignani_modular_2016}.
+It is an interesting observation that modular secure compilation, combined with step-bounded reasoning is sufficient to cover dynamic compromise scenarios without the need to build in dynamic compromise scenarios into the definition of secure compilation like \citet{Abate:2018:GCG:3243734.3243745}.
 

### 2020-12-14 popl-2019-journal-v/proving-full-abstraction.tex
Add file with JFP reviews in order to collect responses, fix trivial typos already.

@@
 The invariants of a world can vary in complexity and expressiveness depending on the application.
-The possible contents of the memory also influences what the world looks like.
+The possible contents of the memory also influence what the world looks like.
 For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds using a form of memory ownership assumptions.
@@
 We cannot let multiple linear capabilities reference the same memory.
-The spatial region represent ownership of a part of memory which ensures that a linear capability is not aliased.
-Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
+The spatial region represents ownership of a part of memory which ensures that a linear capability is not aliased.
+Even though spatial regions give the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
 To this end, we have shadow regions.
@@
 However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
-Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
+Instead, the stack frame and code return point are connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
 In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
@@
 The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
-The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities \emph{uniquely} references part of memory.
-
-Given a world, we want to be able to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
+The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities \emph{uniquely} reference part of memory.
+
+Given a world, we want to be able to express that a capability that is otherwise valid with respect to the world is not linear or indirectly depends on a linear capability.
 This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions and then requiring that the capability is valid w.r.t.\ this new world.
@@
 
-Safety of capabilities should be defined, so it is monotone with respect to the future world relation.
+Safety of capabilities should be defined, so that it is monotone with respect to the future world relation.
 This means that the system assumptions that capabilities rely on for safety are defined such that capabilities remain safe during execution.
@@
 The $\lrstk[]$ relation also partitions the \trgcm{} memory but only the \trgcm{} as the \srccm{} partitions are given as the stack frames.
-The stack relation also partitions the authority of the world, so it can relate the stack frames in a way that respect linearity.
+The stack relation also partitions the authority of the world, so it can relate the stack frames in a way that respects linearity.
 The stack on \srccm{} represents the call stack which means that each stack frame corresponds to a call and its local data.
@@
 We define a \srccm{} configuration to logically approximate a \trgcm{} configuration when the halting termination of the \srccm{} configuration implies the halting termination of the \trgcm{} configuration.
-This also means that \srccm{} configurations that terminates by failing termination are related to any \trgcm{} configuration.
+This also means that \srccm{} configurations that terminate by failing termination are related to any \trgcm{} configuration.
 Intuitively, this is because the $\failed$ configuration signals that there was an attempt to break the guarantees of the capability machine.
@@
 In Figure~\ref{fig:value-relation}, we have sketched the two value relations.
-This shows that for the most part, words on \srccm{} are related to words on \trgcm{} that are syntactical identical.
+This shows that for the most part, words on \srccm{} are related to words on \trgcm{} that are syntactically identical.
 The only exception is stack pointers on \srccm{} that are related to linear capabilities on \trgcm{}.
@@
 We use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
-The assumptions we can make on the code depends on whether it is part of a trusted or untrusted component.
+The assumptions we can make on the code depend on whether it is part of a trusted or untrusted component.
 This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
@@
 The code capability for a closure must point to the code memory because it is the only part of memory that is executable.
-Untrusted components cannot safely have a capability for a trusted components code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
+Untrusted components cannot safely have a capability for a trusted component's code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
 While it is not safe to give a bare capability for a trusted components code memory, it can be perfectly safe to give a sealed capability for a trusted components code.
@@
 Specifically, linear capabilities can only address memory governed by a $\spatialo$ region, and normal capabilities can only address memory governed by a $\pure$ region.
-All the permission based condition we define project the regions that the capability may address from the world.
+All the permission based conditions we define project the regions that the capability may address from the world.
 The addressable $\addressable{}$ function takes care of the projection:
@@
 The $\readCond[]{}$ is compatible with all the operations that can be performed on capabilities.
-This means that if two capabilities, for which $\readCond[]{}$ holds, are spliced together, we can establish that $\readCond[]{}$ holds for the resulting capability.
+This means that if two capabilities for which $\readCond[]{}$ holds are spliced together, we can establish that $\readCond[]{}$ holds for the resulting capability.
 To support this, we require the presence of a set of regions $S$ that governs the addresses the capability has authority over rather than just a single region.
@@
 On the other hand, a $\spatialo$ region can only be used to establish the $\readCond[]{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
-None the less, we need to support arbitrary splitting of linear capabilities, which means that $\readCond[]{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond[]{}$.
-This is why, $\readCond[]{}$ requires all regions to only govern one address when the capability is linear.
+Nonetheless, we need to support arbitrary splitting of linear capabilities, which means that $\readCond[]{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond[]{}$.
+This is why $\readCond[]{}$ requires all regions to only govern one address when the capability is linear.
 This means that after a split, the authority of the regions for the bottom half of the split can go to one capability and the remaining regions can go to the top half.
@@
 A capability with write permission can be used to write to memory.
-The question is, what can we safely allow to be written to memory without any memory invariants being broken.
+The question is what can we safely allow to be written to memory without any memory invariants being broken.
 The answer to this is anything - even words that are unsafe.
@@
 A safe write capability must respect the memory invariant of the region that governs the memory the capability gives access to.
-Now consider the case, where the invariant permits two memory segments that differs in two or more addresses.
+Now consider the case where the invariant permits two memory segments that differ in two or more addresses.
 In this case, the write capability cannot be used to transform the memory from one memory segment to the other because only one memory address can be updated at a time.
@@
 In other words, the adversarial code should be able to transform the memory segment to any memory segment permitted by the region.
-This is captured by address stratification (Definition~\ref{def:address-stratified}) which basically says that if a region permits two memory segments, then all the intermediate memory segments you may end up with when transforming one memory segment to the other must be permitted as will.
+This is captured by address stratification (Definition~\ref{def:address-stratified}) which basically says that if a region permits two memory segments, then all the intermediate memory segments you may end up with when transforming one memory segment to the other must be permitted as well.
 \begin{definition}
@@
 
-The $\readCond[]{}$ and $\writeCond[]{}$ specifically uses the heap sub-world which means that it can only be used for heap capabilities.
+The $\readCond[]{}$ and $\writeCond[]{}$ specifically use the heap sub-world which means that it can only be used for heap capabilities.
 This means that we cannot use it for stack capabilities.
-To take care of stack capabilities, we define two more conditions a $\stackReadCond[]{}$ and $\stackWriteCond[]{}$.
-The two new condition are essentially the same as the $\readCond[]{}$ and $\writeCond[]{}$ except that they use the free stack sub-world and assume that the capability is linear as all stack capabilities are linear.
+To take care of stack capabilities, we define two more conditions: a $\stackReadCond[]{}$ and $\stackWriteCond[]{}$.
+The two new conditions are essentially the same as the $\readCond[]{}$ and $\writeCond[]{}$ except that they use the free stack sub-world and assume that the capability is linear as all stack capabilities are linear.
 Note that we do not have any condition that talks about the stack-frames sub world because we should never have a capability that allows us to directly read from or write to that part of memory.
@@
 \end{definition}
-The final permission, we define conditions for is the execute permission.
+The final permission we define conditions for is the execute permission.
 We define two conditions $\execCond[]{}$ and $\xReadCond[]{}$.
@@
 However, we do not need to consider arbitrary pairs: given a sealed capability we only have to consider the capabilities permitted by the relevant seal invariant.
-Just like the $\lrename$ relation captures what it means for a word to be safe as a program counter, we later define $\lrexjname$ that define what it means for a code and data capability pair to be safe together as program counter and data capability, respectively.
+Just like the $\lrename$ relation captures what it means for a word to be safe as a program counter, we later define $\lrexjname$ that defines what it means for a code and data capability pair to be safe together as program counter and data capability, respectively.
 A sealed capability is safe when it can be paired with any sealed capability from the seal invariant such that the pair is in the $\lrexjname$ relation.
@@
 \label{par:untrusted-val-rel}
-The untrusted value relation $\lrvname_\untrusted$ relates all the words that untrusted components can safely posses.
+The untrusted value relation $\lrvname_\untrusted$ relates all the words that untrusted components can safely possess.
 That is words that cannot be used to break any memory invariants.
@@
 
-For sets of seals to be related related in the untrusted relation they must be syntactically equal.
+For sets of seals to be related in the untrusted relation they must be syntactically equal.
 Further, the seals in the set should be disjoint from the return seals and trusted closure seals ($\gsigrets$ and $\gsigcloss$) because the trusted code relies on having the sole access to them.
@@
 % Code pointers for trusted code
-Two capabilities are related as trusted code pointers if they are normal, has a permission derivable from read-execute, and has a range of authority within the trusted address space, $\ta$.
+Two capabilities are related as trusted code pointers if they are normal, have a permission derivable from read-execute, and have a range of authority within the trusted address space, $\ta$.
 Further, we need to know that the capabilities actually point to a piece of code which is why we require the $\xReadCond[]{}$ to be satisfied.
@@
 
-Executions are related when the observable effect of the executions are permissible.
+Executions are related when the observable effects of the executions are permissible.
 The permissible observations are defined by the observation relation, so we define the expression relation in terms of the observation relation.
@@
 It takes related memories and register files (ignoring the $\rdata$ register) and combines them into two configurations.
-Each of the configurations are plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lroname$ relation.
+Each of the configurations is plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lroname$ relation.
 The $\lrename_\mathrm{xjmp}$ relation is defined in \figurename~\ref{fig:expr-rels}.
@@
 % Include the actual FTLR
-An important lemma in our proof of full abstraction of the embedding of \srccm{} into \trgcm{}, is the fundamental theorem of logical relations (FTLR).
+An important lemma in our proof of full abstraction of the embedding of \srccm{} into \trgcm{} is the fundamental theorem of logical relations (FTLR).
 The name indicates that it is an instance of a general pattern in logical relations proofs, but is otherwise unimportant.

### 2020-12-16 popl-2019-journal-v/paper.tex
Address a few more minor remarks.

@@
 % Paper specific redefinitions of commands
-\renewcommand{\MemSeg}{\shareddom{MemSeg}}
+\renewcommand{\MemFrag}{\shareddom{MemFrag}}
 \renewcommand{\Reg}{\shareddom{RegFile}}
@@
 \todo[inline]{TODOS ENABLED!!!}
-Secure compilation is an active topic of research (e.g.~\citep{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016, patrignani_2019,barthe_formal_2019}), but real secure compilers are still a rare sight.
+Secure compilation is an active topic of research \citep[see, e.g.,~][]{devriese_modular_2017,patrignani_hyper_2017,Abate:2018:GCG:3243734.3243745,new_universal_embedding_2016,juglaret_beyond_2016, patrignani_2019,barthe_formal_2019}, but real secure compilers are still a rare sight.
 Secure compilers preserve source-language (security-relevant) properties even when the compiled code interacts with arbitrary target-language components.
@@
 Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
-Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson2018ISA}.
+Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson_capability_2020}.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 It starts from the premise that security results for a calling convention should be reusable as part of a larger proof of a secure compiler.
-To this end, we define a second operational semantics for our target language with a native well-bracketed call stack and primitive ways to do calls and returns.
-This well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
+To this end, we define two operational semantics for our target language: the first one is unsurprising with just a register bank and a linear memory, but the second features a native well-bracketed call stack and primitive ways to do calls and returns.
+This second, well-behaved semantics guarantees WBCF and LSE natively for components using our calling convention.
 As such, these components can be sure that they will only ever interact with other well-behaved components that respect our desired properties.
@@
 In other words, the authority the underlying capability represents is encapsulated under the seal.
-In order to seal a sealable with a seal $\sigma$, it is necessary to have the authority to do so.
-The permission to make sealed capabilities is represented by a set of seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
+In order to seal a capability with a seal $\sigma$, it is necessary to have the authority to do so.
+The permission to make sealed capabilities is represented by a second form of capability in addition to the memory capabilities we saw above: a set-of-seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
 Such a capability represents the authority to seal other capabilities with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
-In the spirit of memory capabilities, a set of seals capability has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
+In the spirit of memory capabilities, a set-of-seals capability has a current seal $\sigma_{\var{current}}$ that is selected for use in the next seal operation.
 % On \trgcm{}, we represent seals in sets rather than single seals to have a more compact representation of multiple seals and make it possible to have seal allocation (not described here).
@@
 The instruction will be explained in more detail below, but essentially, it unseals the pair of capabilities, transfers control to one of them (the code part of the pair) and makes the other one (the data part) available to the invoked code.
-The combination of sealed capabilities and \texttt{xjmp} gives (1) and (2).
+The combination of sealed capabilities and \texttt{xjmp} gives us the properties (1) and (2) mentioned above: encapsulation of components' authority such that authority is restored upon invocation.
 \begin{figure}[tb]
@@
     \mem \in&\Mem & \defeq & \Addr \fun \Word\\
-    \ms \in  &\MemSeg & \defeq & \Addr \parfun \Word\\
+    \ms \in  &\MemFrag & \defeq & \Addr \parfun \Word\\
     \Phi \in & \ExecConf & \defeq & \Mem \times \Reg\\
@@
 Words on \trgcm{} are either capabilities or data (represented by integers $\ints$).
+% Instructions are stored in memory as integers and can be converted back and forth by unspecified $\encInstr{}$ and $\decInstr{}$ functions.
 We assume a finite set of register names $\RegName$ containing at least the registers $\pcreg$, $\rretd$, $\rretc$, $\rstk$, $\rdata$, $\rtmp{1}$, and $\rtmp{2}$.
 We define register files as functions from register names to words.
-Complete memories map all addresses to words and memory segments map some addresses to words (i.e.\ they are partial functions).
+Complete memories map all addresses to words and memory fragments map some addresses to words (i.e.\ they are partial functions).
 $\trgcm{}$ has two terminated configurations $\halted$ and $\failed$ that respectively signify a successful execution and an execution where something went wrong, e.g.,\ an out-of-bounds memory access.
@@
   \cca{r}{\rn}                                 & \arraycolsep=0pt\array[t]{rl}\updPcAddr{}&(\\
-  &\Phi\updReg{r}{w})\endarray & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = (\sigma_\baddr,\sigma_\eaddr,\sigma)$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = (\sigma_\baddr,\sigma_\eaddr,\sigma+n)$, respectively \\
+  &\Phi\updReg{r}{w})\endarray & $\Phi(\rn) = n \in \ints$ and either $\Phi(r) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$ or $\Phi(r) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$ and $w = ((\perm,\lin),\baddr,\eaddr,\aaddr + n)$ or $w = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma+n}$, respectively \\
     \hline
@@
 The simplest of the capability instructions inspect the properties of capabilities: type (\texttt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
-The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
+The current address (or seal) of a capability (or set-of-seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).
 The \texttt{restrict} instruction reduces the permission of a capability according to the permission order $\le$.
@@
 % geta, cca
-The instruction \texttt{geta} projects the current address (or seal) from a capability (or set of seals), and returns $-1$ for data and sealed capabilities.
-\texttt{cca} (change current address) changes the current address or seal of a capability or set of seals, respectively, by a given offset.
+The instruction \texttt{geta} projects the current address (or seal) from a capability (or set-of-seals), and returns $-1$ for data and sealed capabilities.
+\texttt{cca} (change current address) changes the current address or seal of a capability or set-of-seals, respectively, by a given offset.
 Note that this instruction does not need to use $\linCons{}$ like the previous ones, because it modifies the capability in-place, i.e.\ the source register is also the target register.
@@
 % cseal
-Given a sealable $\vsc$ and a set of seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
+Given a sealable $\vsc$ and a set-of-seals where the current seal $\sigma$ is within the range of available seals, the \texttt{cseal} instruction seals $\vsc$ with $\sigma$.
 % xjmp
@@
 % xjmp to unseal
-\trgcm{} does not have an instruction for unsealing capabilities directly, but it can be (partially) simulated using \texttt{xjmp}.
+\trgcm{} does not have an instruction for unsealing capabilities directly, but we will explain below how it can be (partially) simulated using \texttt{xjmp}.
 
@@
   The latter instruction compares two numbers and writes 1 or 0 to a target register depending on whether one number is less than the other.
-  The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set of seals) to the base address of the range of authority (or range of seals).
+  The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set-of-seals) to the base address of the range of authority (or range of seals).
   This instruction makes it easy to work relatively to the base address of a capability.
@@
   It is worth pointing out that \trgcm{} does not have a direct unsealing instruction to extract a capability from its sealed version when the seal is available, but one can be emulated.
-  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set of seals capability $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
+  Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set-of-seals capability $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
   It is possible to extract $w$ in $\rdata$ using $\txjmp{}{}$ in the following way.
@@
   This does not work for sealed executable capabilities because $\txjmp{}{}$ fails if the data capability is executable.
-  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.
+  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.\footnote{Alternatively, we could remove the current restriction in \texttt{xjmp} that requires the data capability to be non-executable but then teh invoked code would often have to manually perform this check instead.}
 
@@
 % Structure
-A base component $\comp_0$ consists of a code memory segment, a data memory segment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
+A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
 The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
@@
   % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
-  % Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
+  % Variable $\offpc$ is the offset from the first instruction of the call to the set-of-seals, and $\offsigma$ is the offset to the seal within the set.
   % Instructions 1-3: ensure that the stack is non-empty.
@@
   % Instructions 25-26: pop 42 from the stack.
-  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set of seals it uses and $\offsigma$ the offset in the set of seals to the call seal.
+  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set-of-seals it uses and $\offsigma$ the offset in the set-of-seals to the call seal.
   $\stkb$ is the globally agreed on stack base.
@@
       \begin{array}{lcrclcr}
-        \src{\StkFrame} & \defeq & \src{\Addr \times \MemSeg} & \phantom{skipskipsip} & \src{\Stack} & \defeq & \src{ \StkFrame^*}
+        \src{\StkFrame} & \defeq & \src{\Addr \times \MemFrag} & \phantom{skipskipsip} & \src{\Stack} & \defeq & \src{ \StkFrame^*}
       \end{array}
                                                                                                                 }\\
-      \src{\ExecConf} & \defeq & \Mem \times \Reg \; \src{\times \; \Stack \times \MemSeg} \\
+      \src{\ExecConf} & \defeq & \Mem \times \Reg \; \src{\times \; \Stack \times \MemFrag} \\
     \end{array}
@@
 The $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$ judgement has two rules.
-The \textsc{C-Seal} judgement requires address $a$ in $\mscode$ to contain a set of seals that is a subset of the closure seals and the return seals.
+The \textsc{C-Seal} judgement requires address $a$ in $\mscode$ to contain a set-of-seals that is a subset of the closure seals and the return seals.
 \textsc{C-Instr} requires address $a$ to contain an integer, which may be interpreted as instruction.
-If address $a$ is the first address of a series of instructions that may be interpreted as a call, and all the addresses are within the trusted address space, then there must be a set of seals available at the address the call in question expects it and that set of seals must contain the return seal that corresponds to the return seal of this call.
+If address $a$ is the first address of a series of instructions that may be interpreted as a call, and all the addresses are within the trusted address space, then there must be a set-of-seals available at the address the call in question expects it and that set-of-seals must contain the return seal that corresponds to the return seal of this call.
 The return seal in question must be in the $\sigrets[\mathrm{owned}]$ to ensure that no other call uses it.
@@
   \label{def:no-hidden-calls}
-  We say that a memory segment $\mscode$ has no hidden calls iff
+  We say that a memory fragment $\mscode$ has no hidden calls iff
   \[
@@
 Definition~\ref{def:no-hidden-calls} says that if code memory contains an instruction that may be part of a call, then either it is a part of a call or there is some other instruction that witnesses the fact that it is not part of a call.
-Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set of seals available.
+Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set-of-seals available.
 
@@
 This is why the word reasonability is defined relative to the memory and stack.
-For instance, if a read capability gives access to a piece of memory that contains a set of seals with a return seal, then it would not be reasonable.
+For instance, if a read capability gives access to a piece of memory that contains a set-of-seals with a return seal, then it would not be reasonable.
 This is why stack pointers and memory capabilities with read permission must only give access to memory with reasonable words.

### 2020-12-16 fully-abstract-comp/technical-report.tex
Fix a lot of R1 remarks.

@@
 %   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-%             { (\pure,H,H_\sigma) \nsubsim (v',H') }
+%             { (\pure,H,H_{\mathrm{seal}}) \nsubsim (v',H') }
 % \and
 %   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-%             { (v,H) \nsubsim (\pure,H',H_\sigma) }
+%             { (v,H) \nsubsim (\pure,H',H_{\mathrm{seal}}) }
 % \and
 %   \inferrule{ \forall \hat{W} \ldotp H \; \hat{W} \nsubsim H' \; \hat{W}}
-%             { (\pure,H,H_\sigma) \nsubsim (\pure,H',H_\sigma') }
+%             { (\pure,H,H_{\mathrm{seal}}) \nsubsim (\pure,H',H_{\mathrm{seal}}') }
 % \end{mathpar}
@@
 \begin{align*}
-  (\pure,H,H_\sigma) \oplus (\pure,H,H_\sigma) =  & \; (\pure,H,H_\sigma) \\
+  (\pure,H,H_{\mathrm{seal}}) \oplus (\pure,H,H_{\mathrm{seal}}) =  & \; (\pure,H,H_{\mathrm{seal}}) \\
   (\spatial,H) \oplus (\spatial,H) =  & \; (\spatial,H) \\
@@
       \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
-      \quad \forall j < i \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
+      \quad \forall j < i \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp
+      a < a' < \stkb) \wedge\\
       \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
@@
       \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
-      \quad \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
+      \quad \dom(\pwheap(r).H_{\mathrm{seal}}) = R_\var{seal}(r)
     \end{array}
@@
       (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
-      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
+      \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_{\mathrm{seal}}) \tand \\
+      \quad H_{\mathrm{seal}} \; \sigma \nequal H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
       \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
       \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
@@
       \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+      \quad \pwheap(r) = (\pure,\_,H_{\mathrm{seal}}) \tand H_{\mathrm{seal}} \; \sigma' \nequal (\lrv \circ \xi)
     \end{array}
@@
 H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc,
-H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+H^\mathrm{code}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \; \gc)
 \]
@@
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \;
   (\ta,\stkb) \; \sigma \; \hat{W}= \\
@@
 %   \begin{equation*}
-%     \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma \not\nequal (\lrv \circ \xi)
+%     \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_{\mathrm{seal}}~ \sigma \not\nequal (\lrv \circ \xi)
 %   \end{equation*}
@@
 
-%     % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma~\hat{W}$.
+%     % In this case, we can choose a $b,e$ and construct a $\hat{W}$ such that $\npair[0]{\left(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1)\right)}$ is not in $\lrv(\xi(\hat{W}))$, but it is in $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_{\mathrm{seal}}~ \sigma~\hat{W}$.
 %     % It suffices to make choices for $b,e$ and $\hat{W}$ such that the following hold:
@@
 %     %  DOMI: what if code has no call or no seal?
-%     It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_\sigma~ \sigma~ \hat{W}$.
+%     It suffices to observe that $\npair[0]{(0,0)} \in \lrv(\xi(\hat{W}))$ but not $\npair[0]{(0,0)} \in \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb,\gsigrets,\gsigcloss)}.H_{\mathrm{seal}}~ \sigma~ \hat{W}$.
 
@@
 
-    By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets,\sigcloss,\mscode$ such that $\pwheap(r) = (\pure,\_,H_\sigma)$ and $H_\sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$, so in particular for $n' = n-1$.
-    It follows easily from the above that also $\npair[n-1]{(\vsc,\_)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \; \xi^{-1}(W)$.
+    By $\npair{(w,\_)} \in \lrv(W_w)$, we get some region $r \in \dom(\pwheap)$ and a set of return seals, a set of closure seals and a code memory: $\sigrets,\sigcloss,\mscode$ such that $\pwheap(r) = (\pure,\_,H_{\mathrm{seal}})$ and $H_{\mathrm{seal}} \nequal H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \mscode \; \gc)$ and $\npair[n']{(\sealed{\sigma,\vsc},\_)} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W)$ for all $n' < n$, so in particular for $n' = n-1$.
+    It follows easily from the above that also $\npair[n-1]{(\vsc,\_)} \in H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \; \xi^{-1}(W)$.
 
@@
   \begin{equation*}
-    H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \nequal \lrv
+    H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \; (\ta,\stkb) \; \sigma \nequal \lrv
   \end{equation*}
@@
   \end{itemize}
-  then $\codereg{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma \sigma \nequal \lrv$ for all $n$
+  then $\codereg{\sigrets,\sigcloss,\mscode,\gc}.H_{\mathrm{seal}} \sigma \nequal \lrv$ for all $n$
 \end{lemma}
 \begin{proof}
-  The result follows easily by definition of $\codereg{}$ and $H^\mathrm{code,\square}_\sigma$.
+  The result follows easily by definition of $\codereg{}$ and $H^\mathrm{code,\square}_{\mathrm{seal}}$.
 \end{proof}
@@
       \begin{equation*}
-        \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal \lrv
+        \forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \pwheap(r) = (\pure,\_,H_{\mathrm{seal}}) \tand H_{\mathrm{seal}} \; \sigma' \nequal \lrv
       \end{equation*}
@@
     \begin{itemize}
-    \item $\npair[n']{(\vsc,\vsc)} \in H_\sigma \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
-
-     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that
+    \item $\npair[n']{(\vsc,\vsc)} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
+
+     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_{\mathrm{seal}}$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that
      one of the following holds: 
@@
    \item If $\isLinear{\vsc}$, then 
-     for all $W' \future W$, $W_o$, $n' < n$ and $\npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that 
+     for all $W' \future W$, $W_o$, $n' < n$ and $\npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o)$, we have that 
      \begin{equation*}
@@
 
-     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, we know that
+     By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_{\mathrm{seal}}$ and because we know that $\sigma\in\sigcloss$, we know that
      one of the following holds:
@@
    \item 
-       If $\nonLinear{\src{\vsc_S}}$ then for all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_\sigma \; \sigma \; \xi^{-1}(W_o)$, we have that
+       If $\nonLinear{\src{\vsc_S}}$ then for all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o)$, we have that
        \begin{equation*}
@@
     \begin{itemize}
-    \item $\npair[n']{(\vsc,\vsc)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
-
-      By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_\sigma$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that one of the following holds:
+    \item $\npair[n']{(\vsc,\vsc)} \in H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W)$ for all $n' < n$:
+
+      By definition of $\codereg{\sigrets,\sigcloss,\mscode,\gc}$ and $H^\mathrm{code,\square}_{\mathrm{seal}}$ and because we know that $\sigma\in\sigcloss$, it suffices to prove that one of the following holds:
       \begin{itemize}
@@
     \item
-      For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$, we have that
+      For all $W' \future \purePart{W}$, $W_o$, $n' < n$, $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$, we have that
       \begin{equation*}
@@
 
-      Since $\sigma\in\sigcloss$, it follows from $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$ that
+      Since $\sigma\in\sigcloss$, it follows from $\npair[n']{(\vsc_S',\vsc'_T)} \in H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \; \gc \; \sigma \; \xi^{-1}(W_o)$ that
       \begin{equation*}
@@
       \begin{equation*}
-        \dom(\codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H_\sigma) = \sigrets\uplus \sigcloss
+        \dom(\codereg[\mathrm{code},\square]{\sigrets,\sigcloss,\mscode,\gc}.H_{\mathrm{seal}}) = \sigrets\uplus \sigcloss
       \end{equation*}
@@
   \label{lem:unique-h-sigma}
-  If $r,r' \in \dom(\pwheap)$ and $\pwheap(r) = (\pure,\_,H_\sigma)$
-  and $\pwheap(r') = (\pure,\_,H_\sigma')$ and $\sigma \in \dom(H_\sigma)$
-  and $\sigma \in \dom(H_\sigma')$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$,
+  If $r,r' \in \dom(\pwheap)$ and $\pwheap(r) = (\pure,\_,H_{\mathrm{seal}})$
+  and $\pwheap(r') = (\pure,\_,H_{\mathrm{seal}}')$ and $\sigma \in \dom(H_{\mathrm{seal}})$
+  and $\sigma \in \dom(H_{\mathrm{seal}}')$ and $\memSat{\ms_S,\stk,\ms_\stk,\ms_T}{W}$,
   then
   \[
-    H_\sigma = H_\sigma' \text{ and } r = r'
+    H_{\mathrm{seal}} = H_{\mathrm{seal}}' \text{ and } r = r'
   \]

### 2020-12-16 popl-2019-journal-v/discussion.tex
Fix a lot of R1 remarks.

@@
 % - Full abstraction proofs difficult
-Our formulation of WBCF and LSE using a fully abstract overlay semantics has an important advantage with respect to others.
+Our formulation of WBCF and LSE using a fully abstract overlay semantics has an important advantage.
 Imagine that you are implementing a fully abstract compiler for a high-level language, i.e.\ a secure compiler that enforces high-level abstractions when interacting with untrusted target-language components.
@@
   When it comes to fully-abstract compilation, then the answer is that all the properties of the high-level language should be enforced, so the real question is what high-level language we would like.
-  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/CC, then we need to come up with a different enforcement scheme.
+  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \textit{longjmp}, then we need to come up with a different enforcement scheme.
   Further, many high-level languages have exceptions as another form of control-flow not (yet) supported by \stktokens{}.
@@
   % Say something about full-abstractions proofs and feature extensions
-  Full abstraction is a property of the whole language.
+  Full abstraction of a compiler is a property that takes into account the whole source and target languages.
   In other words, a full abstraction proof must consider all features of a language to make sure that the features don't interact in a way that breaks language abstractions.

### 2020-12-16 popl-2019-journal-v/paper.tex
Fix a lot of R1 remarks.

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
+% \usepackage{amsmath}
+% \usepackage{amssymb}
 
@@
 % Paper specific redefinitions of commands
-\renewcommand{\MemFrag}{\shareddom{MemFrag}}
+\newcommand{\MemFrag}{\shareddom{MemFrag}}
 \renewcommand{\Reg}{\shareddom{RegFile}}
@@
 Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
-The $\decInstr{} : \Word \fun \Instr$ should be surjective and injective for all non-$\fail$ instructions.
-Further, it should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
+The $\decInstr{} : \Word \fun \Instr$ should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
 \[
@@
 \]
-The $\encInstr{} : \Instr \fun \ints$ function should be injective, and  $\decInstr{}$ is its left inverse, i.e.
+The $\decInstr{}$ should have a right inverse $\encInstr{} : \Instr \fun \ints$, i.e.
 \[
@@
 
-The machine also assumes decode and encode functions for permissions $\decPerm{} : \Perm \fun \ints$ and $\encPerm{} : \ints \fun \Perm$.
-We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$ and surjective.
-For $\encPerm{}$, we assume it is surjective and that it does not encode anything to the $\tgetp{}{}$ error value -1, i.e.
+The machine also assumes decode and encode functions for permissions $\encPerm{} : \Perm \fun \ints$ and $\decPerm{} : \ints \fun \Perm$.
+We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$.
+For $\encPerm{}$, we assume that it does not encode anything to the $\tgetp{}{}$ error value -1, i.e.
 \[
@@
 
-\subsection{Components, linking, programs, and contexts }
+\subsection{Components, linking, programs, and contexts}
 \label{subsec:components-linking}
@@
 The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
-An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
+An adversarial callee may intentionally leave a bit of space in memory above the caller's stack frame, so that they can later return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
 This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
@@
 The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
-At this point (Figure~\ref{fig:stack-base-abuse-b}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
+At this point (Figure~\ref{fig:stack-base-abuse-d}), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
 
@@
 7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
-8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \rtmp{2}{1}\\
+8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \tcca{\rtmp{2}}{1}\\
 9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
@@
   % Instructions 25-26: pop 42 from the stack.
-  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$ with $\offpc$  the offset from line 1 of the call to the set-of-seals it uses and $\offsigma$ the offset in the set-of-seals to the call seal.
+  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$.
+  The number $\offpc$ is the offset between the memory location where the first instruction of the call is stored to the memory location where the set-of-seals capability is stored that should be used for sealing the return capability.
+  This set-of-seals capability may give access to a range of seals and the offset $\offsigma$ identifies which of these should be used.
   $\stkb$ is the globally agreed on stack base.
@@
          \end{multlined}$
-    $\opc, c_\opc =\var{setupOpc(\Phi.\reg(\pcreg))}$ and
+    $\opc, c_\opc =\var{setupOpc}(\Phi.\reg(\pcreg))$ and
     $\stk' = (\opc,\ms_{\stk,\var{local}}) :: \Phi.\stk$ and 
@@
                         $
-    \begin{multlined}
+    \begin{multlined}[t]
       \Phi.\reg(r_1,r_2) =\\ \sealed{\sigma', c_1}, \sealed{\sigma', c_2}
@@
 
-  The well-formedness judgement imposes a quite rigid structure on components: a component's code memory may contain only data and set-of-seals capabilities.
+  The well-formedness judgement imposes a quite rigid structure on components: a component's code memory may contain only data (including encoded instructions) and set-of-seals capabilities.
   The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities (to the component's data memory, sealed with closure seals).
@@
     \forall a \in \dom(\msdata)\ldotp \dom(\mscode),A_{\mathrm{own}}(a),A_{\mathrm{non-linear}},\sigcloss \vdash_{\mathrm{comp-word}} \msdata(a)\\
-    \overline{\dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}}
+    \forall w_{\mathrm{export}} \in \overline{w_{\mathrm{export}}} \ldotp \dom(\mscode), A_{\mathrm{non-linear}}, \sigrets,\sigcloss \vdash_{\mathrm{comp-export}} w_{\mathrm{export}}
   }{
@@
 The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
-The \textsc{Main} rule requires the component to have an entry point in terms of a main pair from the exports.
+The \textsc{Main} rule requires the component to have an entry point in the form of a pair of capabilities $c_{\mathrm{main},c}, c_{\mathrm{main},d}$, which should be part of the exports.
 Further, the base component should satisfy the well-formedness judgement.
@@
       \begin{itemize}
-      \item All of the following hold:
+      \item There exist $\perm$, $\lin$, $\baddr$, $\eaddr$, $\aaddr$, $\aaddr'$ such that all of the following hold:
         \begin{itemize}
@@
     \inference{
-      c_{\mathrm{main},c},c_{\mathrm{main},d}  = \sealed{\sigma, c_{\mathrm{main},c}', c_{\mathrm{main},d}'} &
+      c_{\mathrm{main},c}  = \sealed{\sigma, c_{\mathrm{main},c}'} & c_{\mathrm{main},d} = \sealed{\sigma, c_{\mathrm{main},d}'} &
       \nonExec{c_{\mathrm{main},d}'}\\
@@
 Essentially, they modify standard full abstraction so that preservation and reflection of contextual equivalence are only guaranteed for components that are {\itshape fully defined}, which means essentially that they do not exhibit undefined behavior in any fully defined context.
+If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
+
 In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
-If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-
 Another interesting relation between our work and that of \citet{Abate:2018:GCG:3243734.3243745}, is that they consider dynamic compromise scenarios, where some components start out as trusted until they are compromised and are treated as adversarial after that.

### 2020-12-16 popl-2019-journal-v/proving-full-abstraction.tex
Fix a lot of R1 remarks.

@@
 \]
-The sub-worlds are partial maps from names $\RegName$ (not to be confused with register names), modelled as natural numbers, to invariants.
+The sub-worlds are partial maps from names $\RegionName$ (not to be confused with register names), modelled as natural numbers, to invariants.
 In order to define what this actually means, we need to be more precise about what we mean by invariants.
 
-We call the invariants regions because, as we will see in Section~\ref{subsubsec:ft-and-revocation}, they turn out to not be invariant.
+The invariants are registered in the worlds as regions.
 A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
@@
 \[
-  \Worldh = \RegName \parfun (\World \fun \Rel{\MemSeg \times \MemSeg})
+  \Worldh = \RegionName \parfun (\World \fun \Rel{\MemSeg \times \MemSeg})
 \]
@@
 \[
-  \Worldh = \RegName \parfun (\Regionh \cup \Regions)
+  \Worldh = \RegionName \parfun (\Regionh \cup \Regions)
 \]
@@
 \[
-  \Worldfs = \RegName \parfun \Regions
+  \Worldfs = \RegionName \parfun \Regions
 \]
@@
 \[
-\Worlds = \RegName \parfun (\Regions \times \Addr)
+\Worlds = \RegionName \parfun (\Regions \times \Addr)
 \]
@@
     \quad\exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp\\
-    \qquad \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r)
+    \qquad \dom(W'.i) \supseteq m_i(\dom(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r)
      \endarray
@@
 \end{definition}
-The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible (e.g., the regions are not overlapping).
-Only when we know that a certain memory satisfies the world (as a memory specification, see Section~\ref{subsubsec:mem-sat}), will we be sure that the world's specifications are actually non-contradictory.
+The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible.
+For example, the resulting world might have several regions that impose conflicting requirements on a single memory location.
+However, when we know that a certain memory satisfies the world (as a memory specification, see Section~\ref{subsubsec:mem-sat}), then we can be sure that the world's specifications are non-contradictory.
 
@@
         \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
-        \quad \forall r \in \dom(\erase{\pwheap}{\pure})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
+        \quad \forall r \in \dom(\erase{\pwheap}{\pure})\ldotp \dom(\pwheap(r).H_{\mathrm{seal}}) = R_\var{seal}(r)
       \end{array}
@@
           \quad\forall i \in \{0,\dots,m\} \ldotp (\dom(\src{\ms_i}) \neq \emptyset \wedge\\
-          \quad\quad \forall j < i \ldotp \forall a \in \dom(\src{\ms_i}) \ldotp \forall a' \in \dom(\src{\ms_j}) \ldotp \stkb < a < a') \wedge\\
+          \quad\quad \forall j < i \ldotp \forall a \in \dom(\src{\ms_i}) \ldotp \forall a' \in \dom(\src{\ms_j}) \ldotp a < a' < \stkb) \wedge\\
           \quad\exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
@@
 The stack relation $\lrstk[]$ is similar to the heap relation in some ways.
-The $\lrstk[]$ relation also partitions the \trgcm{} memory but only the \trgcm{} as the \srccm{} partitions are given as the stack frames.
+The $\lrstk[]$ relation also partitions the \trgcm{} memory but not the \srccm{} memory, since the \srccm{} stack memory is already partitioned into stack frames.
 The stack relation also partitions the authority of the world, so it can relate the stack frames in a way that respects linearity.
@@
 Further, the stack frames should be in the part of the memory we agree to be the stack which means that the addresses should be smaller than $\stkb$.
-Informally, this just means that the stack should be laid out in memory as a downwards growing stack with no addresses below $\stkb$.
+Informally, this just means that the stack should be laid out in memory as a downwards growing stack with no addresses above $\stkb$.
 
@@
 Note that each stack frame corresponds to a trusted call.
-Untrusted calls are not protected which means that untrusted stack frames reside in the free stack memory (this does not mean that the untrusted stack frames are necessarily unprotected; it only means that they are not natively protected by \stktokens{}).
+Untrusted calls are not protected which means that untrusted stack frames reside in the free stack memory.
 This means that the protected stack frames are not necessarily packed tightly in memory, and the memory in between is part of the free stack.
-The free stack in between stack frames may be used by untrusted code for their local stack frames (Untrusted stack frames are not protected by \srccm{}, but this does not prevent untrusted code from securing their own stack frames).
+The free stack in between stack frames may be used by untrusted code for their local stack frames As such, untrusted stack frames are not natively protected by \srccm{}.
+However, this does not prevent untrusted code from securing their own stack frames.
 \figurename~\ref{fig:mem-sketch} sketches this.
@@
 %   where the family of equivalences distributes to the underlying c.o.f.e.'s, i.e.
-given shared regions $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
-%   \[
-%   (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
+given shared regions $(\pure,H,H_{\mathrm{seal}}), (\pure,H',H_{\mathrm{seal}}') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
+%   \[
+%   (\pure,H,H_{\mathrm{seal}}) \nequal (\pure,H',H_{\mathrm{seal}}') \text{ iff } H \nequal_\var{MI} H' \wedge H_{\mathrm{seal}} \nequal_{SI} H_{\mathrm{seal}}'
 %   \]
@@
       % \quad\forall \sigma' \in [\sigma_\baddr,\sigma_\eaddr] \ldotp \exists r \in \dom(\pwheap) \ldotp \\
-      % \quad \pwheap(r) = (\pure,\_,H_\sigma) \tand H_\sigma \; \sigma' \nequal (\lrv \circ \xi)
+      % \quad \pwheap(r) = (\pure,\_,H_{\mathrm{seal}}) \tand H_{\mathrm{seal}} \; \sigma' \nequal (\lrv \circ \xi)
 %    \end{array}
@@
       % \isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T} \tand \\
-      % \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      % \qquad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      % \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
+      % \quad\exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \pwheap(r) = (\pure,\_,H_{\mathrm{seal}}) \tand \\
+      % \qquad H_{\mathrm{seal}} \; \sigma \nequal H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      % \qquad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\tand\\
       % \qquad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
-      % \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      % \qquad\quad\forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       % \qquad \qquad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \tand \dots
       % % \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
-      % % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
+      % % \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o) \ldotp \\
       % % \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
@@
 \[
-  \stdreg{A,\gc}{\pure} \defeq (\pur,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
+  \stdreg{A,\gc}{\pure} \defeq (\pure,H_A^{\mathrm{std},\square} \; \gc, \lambda \_ \; \_ \ldotp \emptyset)
 \]
@@
   \left\{
-      \npair{\stpair{\ms}{\ms}} \middle|
-      \begin{array}{l}
-        \dom(\src{\ms_S}) = \dom(\ms_T) \wedge \\
+      \npair{\stpair{\ms'}{\ms'}} \middle|
+      \begin{array}{l}
+        \stpair{\ms'}{\ms'} = \stpair{\ms}{\ms} \wedge \dom(\src{\ms_S}) = \dom(\ms_T) \wedge \\
         \exists S : \dom(\src{\ms_S}) \fun \World \ldotp \xi(\hat{W}) = \oplus_{\aaddr \in \dom(\ms)} S(\aaddr) \wedge\\
@@
     \begin{array}{l}
+      \exists \trust \ldotp
     \dom(\code) = [\baddr,\eaddr] \wedge ([\baddr - 1, \eaddr + 1] \subseteq \ta
@@
 This means that components must use return and closure seals for their intended purpose for \stktokens{} to work.
-The code region has a seal invariant $H^\mathrm{code,\square}_\sigma$ to guarantee that the return and closure seals of the region are used correctly.
+The code region has a seal invariant $H^\mathrm{code,\square}_{\mathrm{seal}}$ to guarantee that the return and closure seals of the region are used correctly.
 The seal invariant is displayed in Figure~\ref{fig:code-reg-seal-inv}.
@@
   \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \;
   (\ta,\stkb,\_,\gsigrets) \; \sigma \; \hat{W} \defeq \\
@@
         \dom(\code) \subseteq \ta \tand\\
-        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \\
+        \exists r \in \dom(\pwpriv[\xi(\hat{W})]) \ldotp \\
         \quad\pwpriv[\xi(\hat{W})](r) \nequal (\stareg[(\ms_S,\ms_T)]{\spatialo,\square} \; (\ta,\stkb), \aaddr'+\calllen) \tand \\
@@
 \begin{multline*}
-  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
+  H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \;
   (\ta,\stkb,\gsigcloss,\gsigrets) \; \sigma \; \hat{W} \defeq \\
@@
 \[
-  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure, H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc, H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \code \; \gc)
+  \codereg{\sigrets,\sigcloss,\code,\gc} \defeq (\pure, H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \code \; \gc, H^\mathrm{code}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \code \; \gc)
 \]
@@
       \qquad (\lin = \linear \Rightarrow \forall r \in S \ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pure}.H
     \end{array}
@@
 For instance, the invariant may require a flag to stay unchanged.
-We express the fact that a region may be more restrictive by making the standard region $\stdreg{R(r),\gc}{\pur}$, which permits all memory segments with safe words, the upper bound of what a region may require when it governs a memory segment that can be accessed through a safe read capability.
+We express the fact that a region may be more restrictive by making the standard region $\stdreg{R(r),\gc}{\pure}$, which permits all memory segments with safe words, the upper bound of what a region may require when it governs a memory segment that can be accessed through a safe read capability.
 
@@
       \qquad (\lin = \linear \Rightarrow \forall r \in S \ldotp |R(r)|  = 1) \wedge\\
-      \qquad \forall r \in S \ldotp \pwheap(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
+      \qquad \forall r \in S \ldotp \pwheap(r).H \nsupeq \stdreg{R(r),\gc}{\pure}.H \wedge\\
       \qquad \quad \pwheap(r) \text{ is address-stratified}
@@
       \qquad \forall r \in S \ldotp |R(r)| = 1\\
-      \qquad \forall r \in S \ldotp \pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsubeq \stdreg{R(r),\gc}{\pure}.H
     \end{array}
@@
       \qquad \forall r \in S \ldotp |R(r)| = 1 \wedge \\
-      \qquad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \wedge\\
+      \qquad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pure}.H \wedge\\
       \qquad \quad \pwfree(r) \text{ is address-stratified}
@@
 \begin{multline*}
-\sealedCond{W,H_\sigma} =\\
+\sealedCond{W,H_{\mathrm{seal}}} =\\
 \left\{\npair{(\sigma, \src{\vsc_S},\vsc_T)}\middle|
   \begin{array}{l}
-  \forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp\\
+  \forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W_o) \ldotp\\
     \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))
@@
       \exists r \in \dom(\pwheap), \sigrets,\sigcloss,\mscode \ldotp \\
-      \quad\pwheap(r) = (\pure,\_,H_\sigma) \tand \\
-      \quad H_\sigma \; \sigma \nequal H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
-      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
+      \quad\pwheap(r) = (\pure,\_,H_{\mathrm{seal}}) \tand \\
+      \quad H_{\mathrm{seal}} \; \sigma \nequal H^\mathrm{code,\square}_{\mathrm{seal}} \; \sigrets \; \sigcloss \; \mscode \; \gc \; \sigma \tand \\
+      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_{\mathrm{seal}} \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
       \quad \isLinear{\src{\vsc_S}} \Rightarrow
       \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in \\
-      \qquad \qquad \qquad \qquad \sealedCond{W,H_\sigma} \wedge\\
+      \qquad \qquad \qquad \qquad \sealedCond{W,H_{\mathrm{seal}}} \wedge\\
       \quad \nonLinear{\src{\vsc_S}} \Rightarrow
       \npair{(\sigma,\src{\vsc_S},\vsc_T)}\in\\ 
-      \qquad\qquad\sealedCond{\mathit{purePart}(W),H_\sigma}\\
+      \qquad\qquad\sealedCond{\mathit{purePart}(W),H_{\mathrm{seal}}}\\
       % This case was never used (?) so it is not necessary.
@@
 Further, we need to allow trusted components to have capabilities for the trusted code which, again, is something that we cannot allow untrusted components to have.
-The untrusted value relation is defined in Figure~\ref{fig:trusted-val-rel}.
+The trusted value relation is defined in Figure~\ref{fig:trusted-val-rel}.
 
@@
   \label{lem:ftlr-comps}
-  If $\comp$ is a well-formed component, i.e. $\wdjud{\comp}$ and either
-    $\dom(\comp.\mscode) \subseteq \ta$ and $\src{\comp}$ is a reasonable component; or
-    $\dom(\comp.\mscode) \mathrel{\#} \ta$,
-  then there exists a $W$ such that
+  If $\comp$ is a well-formed component in \trgcm{}, i.e. $\wdjud{\comp}$.
+  Then consider the same $\src{\comp}$ in \srccm{}.
+  If either $\dom(\comp.\mscode) \subseteq \ta$ and $\src{\comp}$ is a reasonable component in \srccm{}; or $\dom(\comp.\mscode) \mathrel{\#} \ta$, then there exists a $W$ such that
   $\npair{(\src{\comp},\comp)} \in \lrcomp(W)$.

### 2020-12-16 popl-2019-journal-v/paper.tex
Fix more remarks.

@@
 
+\setlength{\floatsep}{30.0pt plus 2.0pt minus 2.0pt}
+
 \begin{document}
@@
 To enforce these properties, target language security primitives are needed to prevent untrusted code from misbehaving without imposing too much overhead on well-behaved code.
-The virtual-memory based security primitives on commodity processors do not seem sufficiently fine-grained to efficiently support this.
+The security primitives based on virtual memory on commodity processors do not seem sufficiently fine-grained to efficiently support this.
 More suitable security primitives are offered by a type of CPUs known as capability machines \citep{levy_capability-based_1984,watson_cheri_2015}.
@@
 %% Sealing
-The operational side of the sealing in \trgcm{} consists of two instructions: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
+Two instructions manipulate seals in \trgcm{}: \texttt{cseal} for sealing a capability and \texttt{xjmp} for unsealing a pair of capabilities.
 % cseal
@@
 
-The machine also assumes decode and encode functions for permissions $\encPerm{} : \Perm \fun \ints$ and $\decPerm{} : \ints \fun \Perm$.
+The full machine semantics in Appendix~\ref{app:instr-interpretation} also assumes decode and encode functions for permissions $\encPerm{} : \Perm \fun \ints$ and $\decPerm{} : \ints \fun \Perm$.
 We assume the $\decPerm{}$ function to be the left inverse of $\encPerm{}$.
@@
         $\srcalt{\var{xjmpRes}(c_1,c_2,}$
-        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1,r_2}{w_1,w_2}\\
+        $\srcalt{\hspace{.25cm}\left(\arraycolsep=0pt\array[c]{rl}\Phi&\updReg{r_1}{w_1}\\
+            &\updReg{r_2}{w_2}\\
             &\updReg{\rretc}{s_c}\\
@@
 
-In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to scenarios where components only start to exhibit undefined behavior after a number of well-defined steps. 
+In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to components with  undefined behavior by proving secure compilation guarantees for the execution until the undefined behavior happens.
 Another interesting relation between our work and that of \citet{Abate:2018:GCG:3243734.3243745}, is that they consider dynamic compromise scenarios, where some components start out as trusted until they are compromised and are treated as adversarial after that.

### 2020-12-16 popl-2019-journal-v/proving-full-abstraction.tex
Fix more remarks.

@@
 \]
-The sub-worlds are partial maps from names $\RegionName$ (not to be confused with register names), modelled as natural numbers, to invariants.
-In order to define what this actually means, we need to be more precise about what we mean by invariants.
-
-The invariants are registered in the worlds as regions.
+
+The sub-worlds are partial maps from names $\RegionName$ (not to be confused with register names), modelled as natural numbers, to regions, which model a form of invariant.
 A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
@@
 %%% spatial/spatial_owned
-We cannot let multiple linear capabilities reference the same memory.
-The spatial region represents ownership of a part of memory which ensures that a linear capability is not aliased.
-Even though spatial regions give the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
-To this end, we have shadow regions.
-A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
-This does not mean that the memory is necessarily not referenced.
-A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t.
-ownership in Section~\ref{subsubsec:joining-worlds}).
+However, spatial regions are not enough.
+A world which contains a spatial region, represents the assumption of exclusive ownership of the represented memory fragment.
+In other words, a linear capability for this memory may be valid in such a world.
+However, when we do not possess exclusive ownership of a memory fragment, that does not mean we should be entirely unaware of the memory's existence.
+To represent knowledge about existence of a spatial region, without requiring its exclusive ownership, we have $\spatial$ regions.
 Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
@@
 \]
+
+A $\spatial$ region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
+This does not mean that the memory is necessarily not referenced.
+When we discuss compatibility of worlds w.r.t.
+ownership in Section~\ref{subsubsec:joining-worlds}, we will see that a world with a shadow region can be combined with a world that has the same region as $\spatialo$.
+This setup will allow ownership of a memory fragment to reside with linear capabilities which are not themselves stored in the memory fragment but elsewhere in memory or in a register.
+In such a setup, the region representing the memory fragment will exist only as $\spatial$ in the world where the memory fragment's contents is valid, but it will exist as $\spatialo$ in the world where the linear capability owning the fragment is valid.
 
@@
 The future world relation can also be thought of as the model of allowed changes in memory over time.
+We will say that $W'$ is a future world of $W$ if $W' \future W$.
 
@@
 Using these Kripke worlds as assumptions, we can then define when different \srccm{} and \trgcm{} entities are related: values, jump targets, memories, execution configurations, components etc.
-The most important relations are summarised in the following table, where we mention the general form of the relations, what type of things they relate and extra conditions that some of them imply:\\
+The most important relations are summarised in the following table, where we mention the general form of the relations, what type of things they relate and extra conditions that some of them imply:
+We use the square symbol as a meta-variable that represents either left-to-right (if $\square = \preceq$) or right-to-left (if $\square = \succeq$) approximation (see Section~\ref{sec:observation-relation}).
+\\
 % latex hack stolen from https://tex.stackexchange.com/questions/78788/align-equations-over-multiple-tabular-rows
@@
   \hline
+  \hline
   \npair{(\src{w_S},w_T)} & \lrv(W) & values (machine words) & safe to pass to adversarial code\\
+  \hline
   \npair{(\src{w_S},w_T)} & \lrvtrusted(W) & values (machine words) & \\
+  \hline
   \npair{(\src{\reg_S},\reg_T)}  &  \lrr(W) & register files & safe to pass to adversarial code\\
+  \hline
   \npair{\src{\Phi_S},\Phi_T}  &  \lro & execution configurations & \\
+  \hline
   \npair{(\src{w_S},w_T)}  &  \lre(W) & $\tjmp{}$ targets &\\
+  \hline
   \left(\arraycolsep=1pt\array{l}(\src{w_{S,1}},\src{w_{S,2}}),\\(w_{T,1},w_{T,2})\endarray\right)  &  \lrexj(W) & $\txjmp{}{}$ targets &\\
+  \hline
   \multicolumn{2}{|c|}{$\memSat{\src{\ms_S},\src{\stk},\src{\ms_\stk},\ms_T}{W}$} & memory & satisfy the assumptions in $W$\\
@@
 \subsubsection{Observation relation}
+\label{sec:observation-relation}
 The observation relation defines what machine configurations have related and permissible observable effects.
@@
 
-So far, we have talked about the logical relation as though we define one.
-However, we actually define two logical approximations that only differ in the observation relation.
-We define a \srccm{} configuration to logically approximate a \trgcm{} configuration when the halting termination of the \srccm{} configuration implies the halting termination of the \trgcm{} configuration.
-This also means that \srccm{} configurations that terminate by failing termination are related to any \trgcm{} configuration.
+So far, we have talked about the logical relation as though we define only a single one.
+However, we actually define two logical relations that only differ in the observation relation.
+The two relations $\lro[\preceq]$ and $\lro[\succeq]$ represent approximation between \srccm{} and \trgcm{} configurations in both directions.
+The former defines that a \srccm{} configuration logically approximates a \trgcm{} configuration when the halting termination of the \srccm{} configuration implies the halting termination of the \trgcm{} configuration.
+This also means that \srccm{} configurations that terminate by failing are related to any \trgcm{} configuration.
 Intuitively, this is because the $\failed$ configuration signals that there was an attempt to break the guarantees of the capability machine.
@@
 
-We define a \trgcm{} configuration to approximate a \srccm{} configuration in a dual way to the above.
+The second observation relation $\lro[\succeq]$ defines that a \trgcm{} configuration approximates a \srccm{} configuration in a dual way to the above.
 \begin{multline*}

### 2020-12-16 popl-2019-journal-v/discussion.tex
Fix some more comments.

@@
   A compiler is secure when it enforces the properties of high-level languages which begs the question what properties we should enforce.
-  When it comes to fully-abstract compilation, then the answer is that all the properties of the high-level language should be enforced, so the real question is what high-level language we would like.
+  When it comes to fully-abstract compilation, then the answer is that all the (observational) equivalences in the high-level language should be enforced, so the real question is what high-level language we would like.
   \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \textit{longjmp}, then we need to come up with a different enforcement scheme.
@@
 
-\subsection{Practical Applicability}
+\subsection{Linear Capabilities in reality}
 We believe there are good arguments for practical applicability of \stktokens{}.
@@
 
-\dominique{Start new subsection here?}
+\subsection{Support for source language idioms and features}
 
 \begin{jversion}
-  Programming languages with a C-like calling convention often allow programs to pass stack references in calls.
+  Our calling convention supports all source language features that can be expressed in \srccm{}.
+  This includes relatively basic features such as optimized tail calls (see Section~\ref{subsec:overlay-semantics}), stack-allocated variable-size arrays (like with C's alloca).
+  Some other idioms require a bit more thought, such as the control flow primitives we discussed above.
+
+  Also interesting is the idiom of passing stack references in calls, as is common in programming languages with a C-like calling convention.
   \stktokens{} supports stack references but with a couple of caveats.

### 2020-12-16 popl-2019-journal-v/paper.tex
Fix some more comments.

@@
 The non-duplicability of linear capabilities together with some security checks allows us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
+To avoid confusion, it is worth pointing out that our linear caabilities are really affine rather than linear: erasing them is allowed.
+However, we have chosen to use the term ``linear'' to align with common usage in the world of linear type systems.
 
@@
 In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
+The current address is a CHERI design choice that makes it easier to implement pointers as capabilities in C-like languages \cite{woodruff_cheri_2014}.
 The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
@@
 % seals
-A sealed capability $\sealed{\sigma,\vsc}$ is a pair of a seal $\sigma$ and a capability $\vsc$.
+A sealed capability $\sealed{\sigma,\vsc}$ is a pair of a seal $\sigma$ (represented simply as a natural number) and a capability $\vsc$.
 A sealed capability makes an underlying capability opaque which means that the underlying capability cannot be changed or used for the operations it normally gives permission to.

### 2020-12-18 popl-2019-journal-v/discussion.tex
Fix some more issues, improve writing in section 4.2.

@@
 \subsection{Support for source language idioms and features}
+\label{sec:supp-source-lang}
 
@@
   We are unsure whether it is realistic to apply this approach for existing C code.
+
+  \stktokens{} relies on an important property: the fact that return seals should be uniquely associated to a single call-site.
+  If we were to use a single return seal for multiple call-site, that would allow an adversary to combine a valid return data capability with a return code capability for the wrong call-site, thus breaking WBCF guarantees.
+  This property is currently enforced statically in the well-formedness property for components' code memory (see Section~\ref{sec:well-form-reas}), specifically in the \texttt{C-Instr} rule.
+  For any call instruction in trusted components' code memory, it is required that the return seal used is exclusively owned by this call-site.
+
+  To guarantee that this property does not just hold when the system is initialized, but continues to hold during execution, the well-formedness judgement additionally enforces a Write-Xor-Execute policy for components:
+  writable capabilities are not allowed for code memory and executable capabilities are not allowed for data memory.
+  Although this approach works, it has the side effect of disallowing any form of dynamic code generation.
+  This is unfortunate, because we believe that in principle, \stktokens{} is compatible with dynamic code generation, as long as the generated code respects the unique allocation of return seals to call sites.
+
+  To lift this restriction, an alternative approach would be to enforce the unique allocation of return seals to trusted call sites semantically rather than statically.
+  More concretely, the semantic reasonability requirement for components could additionally require that during any execution, a component uses a return seal only for a single call-site.
+  Such an approach would, we think, obviate the need for the Write-Xor-Execute policy and the static enforcement of proper use of return seals.
+  Thus, we think dynamic code generation could be allowed without losing \stktokens{} guarantees.
 \end{jversion}

### 2020-12-18 popl-2019-journal-v/paper.tex
Fix some more issues, improve writing in section 4.2.

@@
   The components introduced in Section~\ref{subsec:components-linking} are pretty much unconstrained.
-  For instance, a component can have multiple linear capabilities for the same piece of memory, and there are no restrictions on seals.
+  For instance, a component can have multiple linear capabilities for the same piece of memory, and might have arbitrary set-of-seals capabilities.
   In a real system, the operating system and linker  % ???
@@
 
-  The component notion is used for both trusted and adversarial components.
-  We expect the linker to link the two programs together making sure that all the system invariants are respected.
-  We model a linker that relies on a certain fixed shape for all components.
-  The proper linker behaviour and the syntactic expectations we have of both trusted and untrusted components are captured in a syntactic well-formedness judgement which we explain in this section.
-
-  The well-formedness judgement imposes a quite rigid structure on components: a component's code memory may contain only data (including encoded instructions) and set-of-seals capabilities.
-  The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities (to the component's data memory, sealed with closure seals).
-  This makes a clear separation between code and data memory as neither can contain capabilities for the other.
-  The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking return seals that were indirectly accessible through the data structure.
-  Informally, a well-formed component respects Write-XOR-Execute which means that all capabilities for code memory at most have read-execute permission and all capabilities for data memory at most have read-write permission.
+  In this section, we introduce a syntactic well-formedness judgement, which captures requirements that would normally be enforced by the operating system and linker. 
+  It applies to both trusted and adversarial components.
+  Well-formedness will only be required when the system starts executing, i.e.\ the requirement serves to ensure a sane initial state.
+  It does not prevent code from organizing its memory differently when the system is executing.
+
+  For simplicity, the well-formedness judgement imposes a quite rigid structure on components.
+  A component's code memory may contain only data (including encoded instructions) and set-of-seals capabilities.
+  The data memory may contain only data, memory capabilities to the component's data memory (respecting linearity) or sealed memory capabilities to the component's data memory, sealed with closure seals.
+  This produces a clear separation between code and data memory as neither can contain capabilities for the other.
+  % The separation of the two kinds of memory means that data structures in data memory can be shared without risking unintentionally leaking return seals that were indirectly accessible through the data structure.
+  Well-formedness does not allow executable capabilities to data memory or writable capabilities to code memory, effectively enforcing a kind of Write-XOR-Execute policy.
+  This means we currently exclude dynamic code generation, but see Section~\ref{sec:supp-source-lang} for thoughts on how this restriction might be lifted.
   A component's exports should be sealed code or data capabilities, sealed with a closure seal.
 
-  As discussed before, we only provide guarantees (LSE and WBCF) for components respecting certain semantic rules (reasonable use of seals).
-  Components that do not satisfy these requirements are still allowed in \srccm{}, but (as we will see below) function calls in such components will not behave according to the special \srccm{} semantics that guarantees LSE and WBCF.
-  To distinguish trusted components (which satisfy reasonability requirements) from adversarial components (which don't), we make use of a set of trusted addresses $\ta$.
-  Any component's code memory must fall entirely within or outside of $\ta$, so every component is either trusted or adversarial.
-  Some well-formedness requirements are specific for adversarial and trusted components respectively.
+  As explained in the previous section, we make use of a set of trusted addresses $\ta$ to distinguish trusted components from adversarial components.
+  To prevent ambiguity, the well-formedness judgement requires that any component's code memory must fall entirely within or outside of $\ta$, so every component is either trusted or adversarial.
+  Additionally, some well-formedness requirements are specific for adversarial and trusted components respectively.
   Well-formed adversarial components cannot have return seals.
@@
 \begin{jversion}
-These requirements are formally expressed by the well-formedness judgement $\wdjud[\ta]{\comp}$, i.e.\ it defines when components satisfy the initial syntactic requirements necessary to be able to rely on unique linear capabilities, component unique seals, etc.
+These requirements are formally expressed by the well-formedness judgement $\wdjud[\ta]{\comp}$: it defines initial syntactic requirements for components, necessary to be able to rely on unique linear capabilities, component unique seals, etc.
 The judgement is defined in \figurename~\ref{fig:well-formed} with auxiliary judgements in \figurename~\ref{fig:well-formed-aux}.
@@
 The $\wdjud[\ta]{\comp}$ judgement has two rules: \textsc{Main} and \textsc{Base}.
-The \textsc{Main} rule requires the component to have an entry point in the form of a pair of capabilities $c_{\mathrm{main},c}, c_{\mathrm{main},d}$, which should be part of the exports.
-Further, the base component should satisfy the well-formedness judgement.
-The \textsc{Base} rule ensures that the component has a certain structure.
-Roughly speaking, the \textsc{Base} rule requires the following:
+If components contain a main entry point (in the form of capabilities $c_{\mathrm{main},c}, c_{\mathrm{main},d}$), then the \textsc{Main} rule requires them to be part of the exports.
+The \textsc{Base} rule for base components has a variety of requirements:
 \begin{itemize}
@@
   This prevents code memories from being spliced together.
-  If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow.
-  Further, the possible control-flows of a component suddenly depend on the memory locations of all the components\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
+  If the capabilities for two code memories can be spliced together, then the execution of one code memory can continue into the other creating an unintended control-flow which would cause various complications\footnote{Alternatively to this syntactic condition on components, we could require trusted components to never splice executable capabilities of unknown origin.}.
 \item \textit{All memory can either be addressed by any number of non-linear capabilities or at most one linear capability.}
-  The data address space is split into $A_\var{non-linear}$ and $A_\var{linear}$ that can be addressed by non-linear capabilities and linear capabilities, respectively.
-  The judgement ensure uniqueness of linear capabilities by letting each address of the data memory take sole ownership of addresses in $A_\var{linear}$.
+  The data address space is split into $A_\var{non-linear}$ and $A_\var{linear}$ which can be addressed by non-linear capabilities and linear capabilities, respectively.
+  The judgement ensures uniqueness of linear capabilities by allocating ownership of addresses in $A_\var{linear}$ uniquely to linear capabilities in data memory.
 \item \textit{All import addresses are part of the data memory.} 
 \item \textit{Import and export symbols are disjoint.}
-\item \textit{One of the following is true}
+\item \textit{One of the following is true:}
   \begin{itemize}
   \item \textit{The code address space is disjoint from the trusted address space $\ta$ and there are no return seals.} In this case, the component contains untrusted code.
-We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals\footnote{Untrusted code still has access to closure seals that could be used to protect calls.}.
-  \item \textit{The code address space is part of the trusted address space $\ta$.} In this case, the component contains trusted code.
-We do not impose requirements on the return seals here (the auxiliary judgements will impose restrictions on them), so the trusted component can have the return seals necessary for its calls.
+We are interested in WBCF and LSE from the perspective of the trusted code, so we do not let untrusted memory have return seals.
+    Untrusted code still has access to closure seals that it can use to protect calls.
+  \item \textit{The code address space is part of the trusted address space $\ta$.}
+    In this case, the component contains trusted code and can use return seals for calls.
   \end{itemize}
-\item \emph{The data address space is disjoint from the trusted address space $\ta$.} Data memory is not executable, so the trusted addresses never include the data memory addresses.
-\item \textit{The code memory satisfies the component-code well-formedness judgement (\figurename~\ref{fig:well-formed-code})}, allowing only seal capabilities for seals owned by the component and numbers.
-If the numbers correspond to the encoding of a call instruction, then they must reference a unique return seal in the code memory.
-\item \textit{Each word in the data memory satisfies the component-word judgement (\figurename~\ref{fig:well-formed-word}).} This must be with respect to the linear addresses assigned to this address.
-It allows only numbers, linear and non-linear capabilities to memory owned by the component and values sealed with closure seals.
-\item \textit{All the exports satisfy the components-export well-formedness judgement (\figurename~\ref{fig:well-formed-export})}, requiring that they are well-formed by the component-word judgement and don't own linear memory.
+\item \textit{The data address space is disjoint from the trusted address space $\ta$.}
+  Data memory is not executable, so the trusted addresses never include data memory addresses.
+\item \textit{The code memory, data memory and exports satisfy, respectively, the component-code, component-word and components-export well-formedness judgements}, which we explain next.
 \end{itemize}
@@
     \sigrets,d_\sigma(a),\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a\\
-    \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
+    % \exists a \ldotp \mscode(a) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \wedge [\sigma_\baddr,\sigma_\eaddr] \neq \emptyset
   }{
@@
 \end{figure}
-\figurename~\ref{fig:well-formed-word} defines the judgement $\vdash_{\mathrm{comp-word}}$ which specifies the words that are well-formed in a component's data memory.
-The judgement is defined by three rules: \textsc{W-Data}, \textsc{W-Capability}, and \textsc{W-Sealed-Capability}.
+\figurename~\ref{fig:well-formed-word} defines the well-formedness judgement $\vdash_{\mathrm{comp-word}}$ for words in a component's data memory.
 The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
@@
 Finally, if it is non-linear, then it must be in the non-linear address space.
-The \textsc{W-Sealed-Capability} specifies that well-formed sealed capabilities in data memory are sealed with a closure seal and whatever is sealed is itself a well-formed word.
+The \textsc{W-Sealed-Capability} allows capabilities sealed with a closure seal if the underlying capability is itself a well-formed word.
 Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
@@
 
-\figurename~\ref{fig:well-formed-export} defines the well-formed exports $\vdash_{\mathrm{comp-export}}$.
-\textsc{E-Word} says that any well-formed word for data memory is well-formed as an export.
-In particular, this means that sealed capabilities for the data memory are acceptable exports which allows components to export the data part of a sealed capability pair.
-The \textsc{E-Sealed-Code} rule allows the code part of a sealed capability pair to be exported.
-In particular, it allows a capability with read-execute permission and its range of authority within the component's code address space to be exported when it has been sealed with a closure seal.
-Together, these two rules allow components to export closures.
-
-\figurename~\ref{fig:well-formed-code} defines the well-formed code memories $\vdash_{\mathrm{comp-code}}$.
-The well-formed code memories are defined by two judgements.
-One judgement considers the entire memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ while the other considers the contents of each code memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
-Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation.
-The well-formedness of code memory depends on whether seals for calls are located in the place we expect them to be located.
+\figurename~\ref{fig:well-formed-export} defines the well-formedness judgement $\vdash_{\mathrm{comp-export}}$ for component exports.
+Essentially, \textsc{E-Word} says that any well-formed word is also allowed as an export.
+In particular, sealed capabilities for data memory are well-formed, which allows components to export the data part of a sealed capability pair.
+Additionally, \textsc{E-Sealed-Code} allows read-execute capabilities to the component's code address space when they have been sealed with a closure seal.
+Together, these two rules allow components to export closures as sealed code-data capability pairs.
+
+Finally, \figurename~\ref{fig:well-formed-code} defines the well-formedness judgement $\vdash_{\mathrm{comp-code}}$ for component's code memory.
+Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation, because it depends on whether seals for calls are located in the place we expect them to be located.
+Therefore, the main judgement considers the entire code memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ and is defined in terms of an auxiliary judgement for an individual memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
 The $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$ judgement has two rules.
-The \textsc{C-Seal} judgement requires address $a$ in $\mscode$ to contain a set-of-seals that is a subset of the closure seals and the return seals.
-\textsc{C-Instr} requires address $a$ to contain an integer, which may be interpreted as instruction.
-If address $a$ is the first address of a series of instructions that may be interpreted as a call, and all the addresses are within the trusted address space, then there must be a set-of-seals available at the address the call in question expects it and that set-of-seals must contain the return seal that corresponds to the return seal of this call.
-The return seal in question must be in the $\sigrets[\mathrm{owned}]$ to ensure that no other call uses it.
-In order to ensure that return seals are at most used by one call, the $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ judgement takes all the available return seals and partitions them over code addresses.
-This means that no other call can use the same seal.
-Further, return seals cannot be used to seal non-return capabilities, so the \textsc{C-Mem} judgement requires the set of return seals to be disjoint from the set of closure seals.
-To limit the amount of corner cases we have to consider, we require the code memory to have no \textit{hidden calls}.
-\lau{Do we actually use the hidden calls in the proofs?
+The \textsc{C-Seals} rule allows set-of-seals capabilities for closure and return seals.
+The \textsc{C-Instr} rule allows integers, including encoded instructions.
+However, if the code memory contains a call instruction in a trusted component (i.e.\ addresses in $\ta$), then there must be a set-of-seals capability available at the expected address, containing the return seal for this call.
+To ensure that return seals are at most used by one call, the $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ judgement partitions all the available return seals over code addresses and \textsc{C-Instr} requires that the seal used by a call is in $\sigrets[\mathrm{owned}]$.
+Further, the \textsc{C-Mem} judgement requires the set of return seals to be disjoint from the set of closure seals, so return seals cannot be used to seal non-return capabilities.
+% domi: I have investigated this and I don't think we use this any more, as pointed out by Lau below.
+% To rule out certain corner cases, we require the code memory to have no \textit{hidden calls}.
+% \lau{Do we actually use the hidden calls in the proofs?
 I vaguely recall this was added to make sure that splicing would not create a call.
 However, we also have the padding which, at least in part, takes care of this.}
-\begin{definition}[No hidden calls]
-  \label{def:no-hidden-calls}
-  We say that a memory fragment $\mscode$ has no hidden calls iff
-  \[
-    \begin{array}{l}
-      \forall \aaddr \in \dom(\mscode) \ldotp \\
-      \quad \forall i \in [0,\calllen - 1] \\
-      \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
-      \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1] \wedge\\
-      \qquad \qquad \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
-      \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
-      \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
-    \end{array}
-  \]
-\end{definition}
-Definition~\ref{def:no-hidden-calls} says that if code memory contains an instruction that may be part of a call, then either it is a part of a call or there is some other instruction that witnesses the fact that it is not part of a call.
-Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set-of-seals available.
+% \begin{definition}[No hidden calls]
+%   \label{def:no-hidden-calls}
+%   We say that a memory fragment $\mscode$ has no hidden calls iff
+%   \[
+%     \begin{array}{l}
+%       \forall \aaddr \in \dom(\mscode) \ldotp \\
+%       \quad \forall i \in [0,\calllen - 1] \\
+%       \qquad \mscode(a + i) = \scall[i]{\offpc,\offsigma}{r_1}{r_2} \Rightarrow \\
+%       \qquad \quad (\dom(\mscode) \supseteq [\aaddr-i,\aaddr + \calllen - i - 1] \wedge\\
+%       \qquad \qquad \mscode([\aaddr-i,\aaddr + \calllen - i -1]) =
+%       \scall[0..\calllen-1]{\offpc,\offsigma}{r_1}{r_2}) \vee\\
+%       \qquad \quad \exists j \in [\aaddr-i,\aaddr + \calllen - i -1] \cap \dom(\mscode) \ldotp \mscode(j) \neq  \scall[j-\aaddr-i]{\offpc,\offsigma}{r_1}{r_2}
+%     \end{array}
+%   \]
+% \end{definition}
+% The definition requires that if code memory contains an instruction that may be part of a call, then either
+% \begin{itemize}
+% \item it is a part of a call which falls entirely within the code memory or 
+% \item there is some other instruction within the code memory, that witnesses the fact that it is not part of a call.
+% \end{itemize}
+% domi: update: no longer needed either, see TR.
+% Finally, the \textsc{C-Mem} judgement requires that the code at least have one non-empty set-of-seals available.
 

### 2020-12-18 popl-2019-journal-v/paper.tex
Pass over section 4.3

@@
 % Reasonability
-The static guarantees given by $\wdjud[\ta]{\comp}$ make sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a trusted component from doing something silly during execution that undermines \stktokens{}.
-In order for \stktokens{} to provide guarantees for a component, we expect it not to shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
+The static guarantees given by $\wdjud[\ta]{\comp}$ make sure that components initially don't undermine the security measures needed for \stktokens{}.
+However, in order for \stktokens{} to provide guarantees for a component, we additionally expect it not to shoot itself in the foot at run-time and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
 More precisely, we expect four things of a reasonable component:
@@
 \item It checks the stack base address before performing a call.
-As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
-\item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
+As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it would often be redundant.
+\item It uses the return seals only for calls and the closure seals in an appropriate way.
+  Specifically, closure seals should only be used to seal executable capabilities for code that behaves reasonably, or for certain forms of non-executable capabilities.
 \item It does not leak return and closure seals (or means to retrieve them).
-This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
+This means that set-of-seals capabilities with return or closure seals cannot be left in registers when transferring control to another module.
 There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
 \item It never stores return and closure seals (or means to get them) to memory.
-By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
+By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them.
 \end{enumerate}
@@
   \subsubsection{Reasonable words}
-  To provide any guarantees, \stktokens{} rely on the program not to leak return seals in any way.
+  To provide any guarantees, \stktokens{} relies on the program not to leak return seals in any way.
   But what does it mean to ``leak'' a return seal?
-  It means that sets of seals that contain return seals as well as any means to obtain such sets cannot be leaked.
+  It means that set-of-seals capabilities that contain return seals as well as any means to obtain such sets cannot be leaked.
   The following definition makes this more precise.
@@
 \end{definition}
-According to the definition, sets of seals that contain return or closure seals are not reasonable.
+The definition rules out set-of-seals capabilities that contain return or closure seals.
 \stktokens{} rely on return seals to be unique in order to work, but it does not rely on closure seals.
-However, it defeats the purpose of the closure seals if they are leaked to malicious code as it allows the malicious code to fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectually allows malicious code to unseal the data capability.
+However, leaking closure seals to malicious code still defeats their purpose as the malicious code can then fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectively allows malicious code to unseal the data capability.
 Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed with.
-
-Whether a capability is reasonable depends on what it gives access to.
-This is why the word reasonability is defined relative to the memory and stack.
-For instance, if a read capability gives access to a piece of memory that contains a set-of-seals with a return seal, then it would not be reasonable.
-This is why stack pointers and memory capabilities with read permission must only give access to memory with reasonable words.
+Capabilities to trusted code memory in $\ta$ are also ruled out, because the code memory may contain set-of-seals capabilities.
+Capabilities that are sealed with capabilities not owned by the component must still be reasonable themselves (since such an untrusted seal provides no protection at all).
+
+Finally, we define that stack or memory capabilities with read permission are reasonable if the memory they give access to contains only reasonable words.
+To accomodate this, reasonability of words is defined relative to the memory and stack.
 The definition is cyclic, but the step-index ensures that it is well-founded.
@@
 While the step relation is defined over configurations, it is the pc that decides what instruction is executed.
-Therefore, it only seems natural to also define when a pc is reasonable.
-Definition~\ref{def:reasonable-pc} roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file are reasonable and the result is a reasonable configuration.
+Therefore, we define when an executable capability is reasonable as a pc.
+The following definition roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file are reasonable and the result is a reasonable configuration.
   \begin{definition}[Reasonable pc]
@@
 In both cases, we require that all of the argument registers contain reasonable words.
-An execution configuration may need to do other operations than calling other code, and seals should no be leaked at any point during execution.
+An execution configuration may need to do other operations than calling other code, and seals should not be leaked at any point during execution.
 For this reason, Item~\ref{item:dont-leak-private} also says that if the next step is not a call or a jump, then the next execution configuration should also be reasonable.
@@
 
-In our result, we assume that adversarial components are well-formed, but not necessarily reasonable. The well-formedness assumption ensures
-that the trusted component can rely on basic security guarantees provided by the capability machine.
+In our result, we assume that adversarial components are well-formed.
+This assumption ensures that the trusted component can rely on basic security guarantees provided by the capability machine.
 For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.

### 2020-12-18 popl-2019-journal-v/paper.tex
Address some more remarks.

@@
 
-  \caption{Possible ways to abuse stack and return capabilities.}
+  \caption{Possible ways to abuse stack and return capabilities.
+    Grey and white frames indicate inactive and active stack frames respectively.
+    The red frame emphasizes stack frames being skipped by an adversarial return.}
   \label{fig:stack-ret-ptr-abuse}
@@
   \end{subfigure}
-  \caption{Abuse of stack and return capabilities prevention.}
+  \caption{Abuse of stack and return capabilities prevention.
+    Grey and white frames indicate inactive and active stack frames respectively.
+    Linear braces indicate memory owned by linear capabilities.  
+  }
 \end{figure}
@@
   \end{subfigure}
-  \caption{Partial return token used to return out of order.}
+  \caption{Partial return token used to return out of order.
+    Grey and white frames indicate inactive and active stack frames respectively.
+    Linear braces indicate memory owned by linear capabilities.
+    The red frame emphasizes a stack frame that is being skipped by an adversarial return.
+  }
   \label{fig:stk-base-abuse}
@@
   The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$.
+  Registers $r_1$ and $r_2$ are assumed to contain the callee closure as a sealed capability pair.
+  The code uses $\rtmp{1}$ and $\rtmp{2}$ as temporary registers.
   The number $\offpc$ is the offset between the memory location where the first instruction of the call is stored to the memory location where the set-of-seals capability is stored that should be used for sealing the return capability.

### 2020-12-19 popl-2019-journal-v/proving-full-abstraction.tex
Start polishing and shortening section 5.

@@
 \begin{jversion}
-Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds.
+These assumptions are explictly represented using (Kripke) worlds.
 To a first approximation, a world is a semantic model of the memory.
@@
 The invariants of a world can vary in complexity and expressiveness depending on the application.
-The possible contents of the memory also influence what the world looks like.
-For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds using a form of memory ownership assumptions.
+% The possible contents of the memory also influence what the world looks like.
+% For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds using a form of memory ownership assumptions.
 In order to relate \trgcm{} and \srccm{}, we model all the features of \srccm{} in the world which means we have to model:
@@
 The three kinds of memory are modelled by having three sub-worlds where each sub-world is its own little world in a traditional sense.
-Linearity is modelled by adding ownership to certain parts of the world that capabilities can take ownership of ensuring that they are the sole reference for that part of memory.
+Linearity is modelled by adding ownership to certain parts of the world which can be claimed
+exclusively by capabilities.
 Memory satisfaction (the relation that decides whether two memories are related in the world) models the call stack by ensuring that the memory is actually shaped like a stack.
@@
 The sub-worlds are partial maps from names $\RegionName$ (not to be confused with register names), modelled as natural numbers, to regions, which model a form of invariant.
-A region describes a collection of related memory segments, so it is simply represented as a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
-Intuitively, we want to be able to say that two memory segments are related when their content is related.
-That is, for every address in the two memory segments, the words that reside there must be related.
-In Section~\ref{subsec:logical-relation}, we define precisely what it means for words to be related, but for now we provide some intuition.
-If we have two integers, then they are related when they are equal.
-If we instead have two capabilities, then they should also be related if they, in some sense, are equal.
-But what does it mean for capabilities to be \textit{equal}?
-Intuitively, it should mean that the capabilities give you the same authority, for instance, if you have two executable capabilities, they should observably perform the same computation.
-As a capability points to a piece of memory, the authority of the capability depends on the contents of that memory.
-In other words to say whether two capabilities are related, we must know the possible contents of the memory.
-The world expresses the possible memory contents, so our regions are world indexed, i.e.
+Intuitvely, a region is simply a relation over memory segments (i.e.\ a relation in $\Rel{\MemSeg \times \MemSeg}$).
+However, this is not sufficient, because the validity of memory contents must often depend on the world itself.
+In other words, our regions must be world indexed, i.e.
 \[
@@
 If we inline $\Worldh$ in $\World$, then we have a circular equation with no solution because the self-reference happens in a negative position.
-Luckily, we can solve circular equations if we move to a different domain.
-For now, we will ignore the problem and return to the issue in Section~\ref{sec:rec-dom-eq}.
+However, in Section~\ref{sec:rec-dom-eq}, we will explain how we can solve the circular equation by moving to a different domain.
 
@@
 % Linearity
-The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
-In order to model this uniqueness, we need to keep track of which parts of memory are uniquely referenced and make sure that only one linear capability references the unique parts.
+The linear capabilities of \srccm{} and \trgcm{} guarantee sole authority over the memory they reference.
+To model this uniqueness, we need to keep track of which parts of memory are uniquely referenced and make sure that only one linear capability references them.
 %% Bookkeeping done by worlds
-We use the world to keep track of what parts of memory must be uniquely referenced by having two kinds of regions: shared and spatial.
-If a memory segment is governed by a shared region, then normal capabilities may reference it.
-On the other hand, if a memory segment is governed by a spatial region, then only a linear capability may reference it.
+We use the world to do this by having two kinds of regions: shared and spatial.
+Memory governed by shared and spatial regions can only be references by non-linear and linear capabilities respectively.
 
@@
 However, spatial regions are not enough.
-A world which contains a spatial region, represents the assumption of exclusive ownership of the represented memory fragment.
-In other words, a linear capability for this memory may be valid in such a world.
-However, when we do not possess exclusive ownership of a memory fragment, that does not mean we should be entirely unaware of the memory's existence.
+A world which contains a spatial region, represents the assumption of exclusive ownership of the corresponding memory.
+In other words, a linear capability for this memory will be valid only in such a world.
+However, that does not mean all other worlds should be entirely unaware of the memory's existence.
 To represent knowledge about existence of a spatial region, without requiring its exclusive ownership, we have $\spatial$ regions.
@@
 A $\spatial$ region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
-This does not mean that the memory is necessarily not referenced.
-When we discuss compatibility of worlds w.r.t.
-ownership in Section~\ref{subsubsec:joining-worlds}, we will see that a world with a shadow region can be combined with a world that has the same region as $\spatialo$.
+We will see in Section~\ref{subsubsec:joining-worlds} that a world with a shadow region is compatible with a world that has the same region as $\spatialo$.
 This setup will allow ownership of a memory fragment to reside with linear capabilities which are not themselves stored in the memory fragment but elsewhere in memory or in a register.
@@
 \]
-\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
+\srccm{} internalizes the \stktokens{} stack, which will only be referenced by linear capabilities, so the two stack regions only need spatial and shadow regions.
 For instance the $\Worldfs$ is defined as
@@
 \]
-$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
-$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
-Conceptually, this means that each of the stack frames is connected with a return point in some code.
-In a traditional C calling convention, the code return point would even be stored in the stack frame.
-However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
-Instead, the stack frame and code return point are connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
-In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
-To this end, each shared region in $\Worlds$ must be paired with a return address:
+$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack itself and specifically the code location for the return point of every frame.
+In a traditional C calling convention, this return location would simply be stored in the stack frame, but \stktokens{} does not require this.
+It is the code part of the return capability pair that references this return location and it is connected to the stack frame by the fact that it is sealed with the same seal as the linear capability that owns the frame: the data part of the return capability pair.
+To ensure and remember that all of this is setup correctly, the world will also keep track of the return location for every frame:
 \[
@@
 \]
-
-% Ensuring linearity
-The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
-The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities \emph{uniquely} reference part of memory.
+It will in fact also remember the seal used, as we will see in the next Section.
 
@@
 \end{definition}
+
 \subsubsection{Seals}
 \label{subsubsec:seals}
-So far, the world represents a collection of assumptions on the memory contents that value correctness may depend on.
-However, value correctness may depend on other assumptions.
+So far, the world represents a collection of assumptions on the memory contents that value relatedness may depend on.
+However, value correctness may also depend on other assumptions.
 Specifically, \stktokens{} has certain assumption on the seals used for return capabilities and closures.
@@
 For example, two related capabilities remain related when an extra invariant is added on unrelated memory, or when a stack frame that it does not reference is dropped.
-In Kripke logical relations, this kind of assumption changes, that do not invalidate the relatedness of values, are modelled by the future world relation $\future$.
+In Kripke logical relations, such changes that do not invalidate the relatedness of values, are modelled by the future world relation $\future$.
 The future world relation can also be thought of as the model of allowed changes in memory over time.
@@
 
-Safety of capabilities should be defined, so that it is monotone with respect to the future world relation.
-This means that the system assumptions that capabilities rely on for safety are defined such that capabilities remain safe during execution.
-The same holds for the memory invariants stored in the world: capabilities stored in memory should remain valid under allowed changes in system assumptions.
-This means we require the world-indexed memory invariants to be monotone in the world.
-In other words, a pair of memory segments that are related now will stay related in future worlds.
-This changes the spatial regions as follows:
+Relatedness of capabilities should be defined, so that it is monotone with respect to the future world relation.
+In other words, relatedness of capabilities should only be predicated on assumptions in the world that are guaranteed to remain valid in future worlds, and the same holds for memory invariants.
+As such, we require the world-indexed memory invariants to be monotone in the world:
 \[
@@
 \]
-We make a similar change to $\Regionh$. The seal invariants are monotone as well.
-
-Kripke future world relations usually allow extending worlds with extra assumptions, or take steps in protocols that the system was designed to support from the start.
-However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on this assumption any more.
+We make a similar change to $\Regionh$. The seal invariants must be monotone as well.
+
+Kripke future world relations usually allow extending worlds with extra assumptions, or take steps in protocols that the system was designed to follow.
+However, in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on this assumption any more.
 Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory and drop or replace the previous assumption.
@@
 
-We add a $\revoked$ tag to the spatial regions $\Regions$ which results in
+We add a $\revoked$ tag to the spatial regions $\Regions$:
 \[
@@
 \]
-The $\spatialo$-region signifies that a linear capability \emph{may} depend on it which means that it cannot be revoked.
+The $\spatialo$-region may be depended on by a linear capability \emph{may} so we cannot allow it to be revoked.
 On the other hand, no capability can depend on a $\spatial$-region, so it can be safely revoked.
-This may be confusing as a $\spatial$-region does not mean that you ``have'' the capability, but we just explained that having a linear capability means that you can repurpose the memory it points to.
-As we will see later, the logical relation splits worlds with respect to their ownership, so each capability can get its own world with unique ownership to depend on.
-A repurposed linear capability gets an entirely new world to depend on.
-The new world constructed for the repurposed linear capability would have a revoked region in place of the $\spatialo$-region, but the new world would not be a future world of the world the linear capability depended on before.
-After all, a future world has to respect old invariants, but a repurposed capability needs to depend on new invariants.
-While the repurposed capability gets a new world to depend on, all other capabilities in the system depend on the same invariants, so they must be valid with respect to worlds that respect the old invariants, i.e.
+% This may be confusing as a $\spatial$-region does not mean that you ``have'' the capability, but we just explained that having a linear capability means that you can repurpose the memory it points to.
+% As we will see later, the logical relation splits worlds with respect to their ownership, so each capability can get its own world with unique ownership to depend on.
+% A repurposed linear capability gets an entirely new world to depend on.
+% The new world constructed for the repurposed linear capability would have a revoked region in place of the $\spatialo$-region, but the new world would not be a future world of the world the linear capability depended on before.
+% After all, a future world has to respect old invariants, but a repurposed capability needs to depend on new invariants.
+% While the repurposed capability gets a new world to depend on, all other capabilities in the system depend on the same invariants, so they must be valid with respect to worlds that respect the old invariants, i.e.
 future worlds.
-Further, the future worlds must be compatible (agree on everything but ownership) with the new world.
-This means that all the $\spatial$-regions must be replaced with $\revoked$ regions which is why $\spatial$ regions can be turned into $\revoked$ regions.
+% Further, the future worlds must be compatible (agree on everything but ownership) with the new world.
+% This means that all the $\spatial$-regions must be replaced with $\revoked$ regions which is why $\spatial$ regions can be turned into $\revoked$ regions.
 
 We define the future world relation in terms of a future region relation which is displayed in Figure~\ref{fig:ft-reg-rel}.
-Apart from being revoked, a region can stay the same, or a $\spatial$ region can become $\spatialo$ which models the affinity of the linear capabilities.
-Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes the linear capabilities affine.
-A linear capability is safe with respect to a world with some $\spatialo$ regions.
-However, this world, and thus the $\spatialo$ regions, is no longer needed if the linear capability is dropped, so other worlds are free to claim the $\spatialo$ region in place of their $\spatial$ region.
+Apart from being revoked, a region can stay the same, or a $\spatial$ region can become $\spatialo$.
+The latter allows us to reassign the $\spatialo$ region to some other world, when the linear capability owning the region is erased (our linear capabilities are actually affine).
 % describe why "Spatial -> spatialo" ~ affine or at least why we have it?
@@
 \end{figure}
-The above repurposing discussion is exemplified in \figurename~\ref{fig:world-example}.
+The above repurposing is exemplified in \figurename~\ref{fig:world-example}.
 The Figure displays two capabilities $c_{\var{normal}}$ and $c_{\var{linear}}$ that are normal and linear, respectively.
@@
 \end{definition}
-The $\uplus$ world join does not guarantee that the result is a sensible world with respect to authority or memory specification.
+The $\uplus$ world join does not guarantee that the result is a sensible world with respect to authority or memory specification: it may contain regions with conflicting requirements for memory.
 In \sectionname~\ref{subsubsec:mem-sat}, we define memory satisfaction which also acts as a well-formedness judgement.
@@
 \end{definition}
-The $\oplus$ operator, like the $\uplus$ operator, does not guarantee that the resulting world is sensible.
-For example, the resulting world might have several regions that impose conflicting requirements on a single memory location.
-However, when we know that a certain memory satisfies the world (as a memory specification, see Section~\ref{subsubsec:mem-sat}), then we can be sure that the world's specifications are non-contradictory.
-
-The $\uplus$ operator is used in the logical relation for components (\sectionname~\ref{subsec:component-rel}) which specifies (among other things) that the world should specify the presence of the component's data memory.
-Linking two components then produces a new component with both components' data memory.
-The linked component is valid in a world that has the combined memory presence specifications, not the combined authority.
-
-Note also that this picture is further complicated by our usage of non-authority-carrying $\spatial$ regions.
+Like the $\uplus$ operator, the $\oplus$ operator does not guarantee that the resulting world is sensible.
+
+% The $\uplus$ operator is used in the logical relation for components (\sectionname~\ref{subsec:component-rel}) which specifies (among other things) that the world should specify the presence of the component's data memory.
+% Linking two components then produces a new component with both components' data memory.
+% The linked component is valid in a world that has the combined memory presence specifications, not the combined authority.
+
+Note that this picture is further complicated by our usage of non-authority-carrying $\spatial$ regions.
 They are really only in a world $W$ as a shadow copy of a $\spatialo$ region in another world $W'$ that $W$ will be combined with.
@@
 The world definition on its own allows the invariants imposed by regions to be overlapping.
-However, to be able to easily determine what invariants a memory segment must respect, we want the invariants of the regions to impose requirements on disjoint parts of the memory.
-The memory satisfaction relation makes sure that this is the case, so in a sense the memory satisfaction also acts as a well-formedness judgement for worlds.
+However, memory satisfaction will allow only a single region to govern every piece of memory.
+It is in this sense that the memory satisfaction also acts as a well-formedness judgement for worlds.
 
@@
 At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory triple, $\ms_S$ (heap), $\ms_\stk$ (free stack), and $\stk$ (stack frames), from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
-The top-level definition of memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
-It also splits the world in three to distribute the authority of the world.
-The three $\ms_T$ partitions are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrhname$, $\lrstk[]$, and $\lrfree[]$, respectively. 
-The $\lrhname$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
-The $\lrhname$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
-The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
-The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the placement in memory of the stack frames have the correct order relative to each other and the base address of the stack.
-Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
-$\lrfree[]$ also makes sure that the base address of the stack is actually part of the free stack.
+% The top-level definition of memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
+% It also splits the world in three to distribute the authority of the world.
+The three parts of source memory are related to parts of $\ms_T$ by the relations $\lrhname$, $\lrstk[]$, and $\lrfree[]$, respectively. 
+We will consider them in this order and discuss $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ last.
+% The $\lrhname$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
+% The $\lrhname$ relation also ensures that the seal invariants associated with each region have invariants on disjoint sets of seals.
+% The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
+% The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the placement in memory of the stack frames have the correct order relative to each other and the base address of the stack.
+% Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
+% $\lrfree[]$ also makes sure that the base address of the stack is actually part of the free stack.
+
 \begin{definition}[Heap relation]
@@
 \end{definition}
+
 Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
 The $\lrhname$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.
-
 To a large extent, the definition of $\lrhname$ is pretty standard.
-$\lrhname$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
+It assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into memory segment pairs each satisfying the invariant of a region.
 The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction partitions the authority of the world using $\oplus$.
 Each of the memory segment pairs must be in the region invariant with respect to a specific world partition which makes sure that uniqueness of linearity of capabilities is respected.
-
 The heap sub-world contains all seal invariants.
 Similar to memory segments, only one seal invariant should impose restrictions on a seal, which $\lrhname$ makes sure is the case.
+
 \begin{definition}[Free stack relation]
@@
 The free stack relation $\lrfree[]$ is in most regards like the heap relation, $\lrhname{}$.
-The free stack relation, partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under part of the world.
+It partitions the \srccm{} and \trgcm{} free stack memory, it partitions the authority of the world, and it requires the memory segment pairs to be related under part of the world.
 For \stktokens{} to work, it should always work on the same stack.
@@
 As the free stack relation relates the stack of \srccm{} with the memory that represents the stack on \trgcm{}, it makes sure that $\stkb$ is the top address of the free stack address space.
+
 \begin{definition}[Stack relation]
@@
 
-According to \stktokens{} (described in \sectionname~\ref{sec:stktokens-explained}), stack frames must be non-empty, so they are distinguishable from a missing stack frame.
-For this reason, $\lrstk[]$ requires every stack frame to be non-empty.
+$\lrstk[]$ requires every stack frame to be non-empty.
+As described in \sectionname~\ref{sec:stktokens-explained}, \stktokens{} requires non-empty stack frames, so a missing frame can be detected.
 Note that each stack frame corresponds to a trusted call.
@@
 This means that the protected stack frames are not necessarily packed tightly in memory, and the memory in between is part of the free stack.
-The free stack in between stack frames may be used by untrusted code for their local stack frames As such, untrusted stack frames are not natively protected by \srccm{}.
 However, this does not prevent untrusted code from securing their own stack frames.

### 2020-12-19 popl-2019-journal-v/paper.tex
Polish and shorten section 4 a bit more.

@@
 
-\lefttitle{\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using \ldots}
+\lefttitle{\stktokens{}: Enforcing WBCF and LSE Using Linear Capabilities}
 \righttitle{Journal of Functional Programming}
@@
 \figurename~\ref{fig:well-formed-word} defines the well-formedness judgement $\vdash_{\mathrm{comp-word}}$ for words in a component's data memory.
-The \textsc{W-Data} rule says all data, i.e.\ integers, are well-formed words.
-The \textsc{W-Capability} rule specifies that all well-formed capabilities in data memory must at most have read and write permission.
-Further, if the capability is linear, then the range of authority must be within the linear address space owned by this address.
-Finally, if it is non-linear, then it must be in the non-linear address space.
-The \textsc{W-Sealed-Capability} allows capabilities sealed with a closure seal if the underlying capability is itself a well-formed word.
-Because of the last requirement, the data memory cannot initially contain sealed code capabilities.
+Well-formed rules are defined as integers, memory capabilities with at most read and write permission and well-formed words sealed with a closure seal.
+The range of authority of linear and non-linear memory capabilities is required to be available in $A_{\mathrm{linear}}$ (where it is exclusively owned) and $A_{\mathrm{non-linear}}$, respectively.
+Note that data memory is not allowed to initially contain sealed code capabilities.
 However, a component is free to place sealed data capabilities in memory during execution.
@@
 \figurename~\ref{fig:well-formed-export} defines the well-formedness judgement $\vdash_{\mathrm{comp-export}}$ for component exports.
-Essentially, \textsc{E-Word} says that any well-formed word is also allowed as an export.
-In particular, sealed capabilities for data memory are well-formed, which allows components to export the data part of a sealed capability pair.
-Additionally, \textsc{E-Sealed-Code} allows read-execute capabilities to the component's code address space when they have been sealed with a closure seal.
-Together, these two rules allow components to export closures as sealed code-data capability pairs.
-
-Finally, \figurename~\ref{fig:well-formed-code} defines the well-formedness judgement $\vdash_{\mathrm{comp-code}}$ for component's code memory.
-Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation, because it depends on whether seals for calls are located in the place we expect them to be located.
-Therefore, the main judgement considers the entire code memory $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ and is defined in terms of an auxiliary judgement for an individual memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
-The $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$ judgement has two rules.
-The \textsc{C-Seals} rule allows set-of-seals capabilities for closure and return seals.
-The \textsc{C-Instr} rule allows integers, including encoded instructions.
-However, if the code memory contains a call instruction in a trusted component (i.e.\ addresses in $\ta$), then there must be a set-of-seals capability available at the expected address, containing the return seal for this call.
-To ensure that return seals are at most used by one call, the $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ judgement partitions all the available return seals over code addresses and \textsc{C-Instr} requires that the seal used by a call is in $\sigrets[\mathrm{owned}]$.
-Further, the \textsc{C-Mem} judgement requires the set of return seals to be disjoint from the set of closure seals, so return seals cannot be used to seal non-return capabilities.
+Essentially, well-formed exports are either well-formed words or sealed read-execute capabilities to the component's code memory, sealed with a closure seal.
+Both cases together allow components to export closures as sealed code-data capability pairs.
+
+Finally, \figurename~\ref{fig:well-formed-code} defines the well-formedness judgement $\vdash_{\mathrm{comp-code}}$ for components' code memory.
+% Unlike data memory and exports, we cannot say whether code memory is well-formed by considering each word in isolation, because it depends on whether seals for calls are located in the place we expect them to be located.
+The judgement $\sigrets,\sigcloss,\ta\vdash_{\mathrm{comp-code}} \mscode$ is defined in terms of an auxiliary judgement for an individual memory address $\sigrets,\sigrets[\mathrm{owned}],\sigcloss,\ta \vdash_{\mathrm{comp-code}} \mscode,a$.
+The rules allow either set-of-seals capabilities for closure and return seals, or integers, including encoded instructions.
+If the code memory contains a call instruction in a trusted component (i.e.\ addresses in $\ta$), then there must be a set-of-seals capability available at the expected address, containing the return seal for this call.
+Additionally, the judgement ensures that return seals are at most used by one call by partitioning
+available return seals in $\sigrets[\mathrm{owned}]$ over code addresses for use in calls.
+Further, return seals must be disjoint from closure seals.
 % domi: I have investigated this and I don't think we use this any more, as pointed out by Lau below.
@@
 It says that if the configuration points to a seal-instruction, then either the instruction is part of a call and uses a return seal or the instruction seals part of a closure and uses a closure seal.
-In order to construct a closure, one must seal a code capability and a data capability.
-If this is an executable capability, then it should be reasonable as a pc.
-On the other hand, if it is not an executable capability, then this must be the data part of the pair, so it should just be a reasonable word.
+In the latter case, the sealed capability must be reasonable as a pc if it is executable or just a reasonable word if it isn't.
 Definition~\ref{def:reasonable-conf} is cyclic through Definition~\ref{def:reasonable-pc}, so both definitions are step-indexed to break the cycle.

### 2020-12-27 popl-2019-journal-v/discussion.tex
Full pass over paper, fix another reviewer remark.

@@
   A compiler is secure when it enforces the properties of high-level languages which begs the question what properties we should enforce.
-  When it comes to fully-abstract compilation, then the answer is that all the (observational) equivalences in the high-level language should be enforced, so the real question is what high-level language we would like.
-  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \textit{longjmp}, then we need to come up with a different enforcement scheme.
+  When it comes to fully-abstract compilation, then the answer is that all properties in the
+  high-level language should be enforced (or at least the equivalences they imply), so the real question is what high-level language we would like.
+  \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \texttt{longjmp}, then we need to come up with a different enforcement scheme.
   Further, many high-level languages have exceptions as another form of control-flow not (yet) supported by \stktokens{}.
@@
   In other words, if we were to define a compiler phase that targets \srccm{}, then we would also have to show that every program it generates is well-formed and reasonable in order to use the full-abstraction result.
-  Without the reasonability constraint, \stktokens{} would have to enforce reasonability instead.
+  Without the reasonability assumption, \stktokens{} would have to enforce reasonability instead.
   That is, \stktokens{} would have to dynamically ensure that no return seals or means to obtain them are passed in calls.
   Such checks would impose a performance overhead and are in principle unnecessary for correct code.
-  We have instead chosen to rely on the compiler to generate reasonable code that never exhibits the unreasonable behaviour.
-  This should be done in a compiler phase where more information about the original program is available, e.g.\ the compilation phase that commits to the low-level translation.
-  Similarly for the syntactic constraints given by well-formedness.
-  The compiler should make sure to generate code that satisfies the well-formedness judgement, so it can be executed by the machine.
+  We have instead chosen to assume that we are given reasonable code that never exhibits the unreasonable behaviour.
+  A compiler phase where more information about the original program is available should be able to
+  ensure the assumption with relative ease.
+  The syntactic requirement of well-formedness is similar, except that it should be ensured by the
+  compiler and linker together.
 
@@
 Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team as well as our own thoughts on the matter.
+We also refer to the latest CHERI ISA report, which contains a concurrently-developed, unimplemented design for linear capabilities, albeit without \texttt{splice} and \texttt{split} \citep[][\S D.7]{watson_capability_2020}.
 As we understand it, the main problems for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
@@
   Our calling convention supports all source language features that can be expressed in \srccm{}.
-  This includes relatively basic features such as optimized tail calls (see Section~\ref{subsec:overlay-semantics}), stack-allocated variable-size arrays (like with C's alloca).
+  This includes relatively basic features such as optimized tail calls (see Section~\ref{subsec:overlay-semantics}) and stack-allocated variable-size arrays.
   Some other idioms require a bit more thought, such as the control flow primitives we discussed above.
@@
   First of all, the stack capability is linear, so all references to the stack have to be linear.
-  This means that the callee has to treat references linearly.
-  Next, like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (which allows them to return). 
+  This means that the callee has to respect this linearity when using the reference.
+  Next, like the stack capability, the stack references must be given back to the caller on return, so they can reconstruct their original stack capability (allowing them to return). 
   Finally, the encapsulated local stack frame should be a continuous piece of memory (because it has to be addressable by a single capability: the data part of the return capability pair).
@@
   \stktokens{} relies on an important property: the fact that return seals should be uniquely associated to a single call-site.
-  If we were to use a single return seal for multiple call-site, that would allow an adversary to combine a valid return data capability with a return code capability for the wrong call-site, thus breaking WBCF guarantees.
+  If we were to use a single return seal for multiple call sites, that would allow an adversary to combine a valid return data capability with a return code capability for the wrong call-site, thus breaking WBCF guarantees.
   This property is currently enforced statically in the well-formedness property for components' code memory (see Section~\ref{sec:well-form-reas}), specifically in the \texttt{C-Instr} rule.
   For any call instruction in trusted components' code memory, it is required that the return seal used is exclusively owned by this call-site.
-
   To guarantee that this property does not just hold when the system is initialized, but continues to hold during execution, the well-formedness judgement additionally enforces a Write-Xor-Execute policy for components:
   writable capabilities are not allowed for code memory and executable capabilities are not allowed for data memory.
-  Although this approach works, it has the side effect of disallowing any form of dynamic code generation.
+  Although this approach works, it has the side effect of disallowing dynamic code generation.
   This is unfortunate, because we believe that in principle, \stktokens{} is compatible with dynamic code generation, as long as the generated code respects the unique allocation of return seals to call sites.

### 2020-12-27 popl-2019-journal-v/paper.tex
Full pass over paper, fix another reviewer remark.

@@
 Instead of CHERI's local capabilities, it builds on {\itshape linear} capabilities; a new form of capabilities that has not previously been described in the published literature, although related ideas have been described by~\citet[``scarce objects'']{szabo_formalizing_1997,szabo_scarce_objects} and in technical documents.
-Concurrently with our work, Watson et al. have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson_capability_2020}.
+Concurrently with our work, \citeauthor{watson_capability_2020} have developed a (more realistic) design for linear capabilities in CHERI that is detailed in the latest CHERI ISA reference~\citep{watson_capability_2020}.
 % KJAA: Perhaps make it clear by who and when/where?
@@
 The non-duplicability of linear capabilities together with some security checks allows us to guarantee WBCF and LSE without large overhead on boundary crossings and in particular without the need for clearing large blocks of memory.
-To avoid confusion, it is worth pointing out that our linear caabilities are really affine rather than linear: erasing them is allowed.
+To avoid confusion, it is worth pointing out that our linear capabilities are really affine rather than linear: erasing them is allowed.
 However, we have chosen to use the term ``linear'' to align with common usage in the world of linear type systems.
@@
 \trgcm{} is mainly inspired by CHERI~\citep{watson_cheri_2015} with linear capabilities as the main addition.
-For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.
+For simplicity, \trgcm{} assumes an infinite address space and unbounded integers.\footnote{Although we have not thoroughly investigated, we do not believe our results depend heavily on these assumptions and we expect they could be lifted without a great impact on the proofs.}
 % \lau{Do we have a good reason to assume an infinite address space for this work?
@@
 The concept of a capability is the cornerstone of any capability machine.
-In its simplest form, a capability is a permission and a range of authority.
-The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
+In its simplest form, a (memory) capability is a permission and a range of authority.
+The former dictates the operations the capability can be used for, and the latter specifies the range of memory it can act upon.
 % Permissions and Range of authority
-The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
+Memory capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}).
 Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
@@
 In order to seal a capability with a seal $\sigma$, it is necessary to have the authority to do so.
-The permission to make sealed capabilities is represented by a second form of capability in addition to the memory capabilities we saw above: a set-of-seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
+The permission to make sealed capabilities is represented by a second form of capability (in addition to the memory capabilities we saw above): a set-of-seals capability $\seal{\sigma_\var{base},\sigma_\var{end},\sigma_{\var{current}}}$.
 Such a capability represents the authority to seal other capabilities with seals in the range $[\sigma_\var{base},\sigma_\var{end}]$.
@@
 The \texttt{store} and \texttt{load} instructions are fairly standard.
-They require a capability with permission to either write or read depending on the operation, they check that the capability points within the range of authority.
+They require a capability with permission to write or read (respectively), they check that the capability points within the range of authority.
 Linear capabilities introduce one extra complication for \texttt{load} as it needs to clear the loaded memory address when it contains a linear capability in order to not duplicate the capability.
@@
 A pair of sealed capabilities can be seen as a closure where the code capability (the capability placed in $\pcreg$) is the program and the data capability is the local environment.
-Because of the opacity of sealed capability, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ sealed with the same seal.
+Because of the opacity of sealed capabilities, the creator of the closure can be sure that execution will start where the code capability points and only in an environment with the related data, i.e.\ with access to a data capability sealed with the same seal.
 This makes \texttt{xjmp} the mechanism on \trgcm{} that transfers control between security domains.
@@
 % split/splice
-Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit what a capability can do before it is passed away.
+Instructions for reducing the authority of capabilities are commonplace on capability machines as they allow us to limit a capability's authority before passing it to other code.
 For normal capabilities, reduction of authority can be done without actually giving up any authority by duplicating the capability first.
-With linear capabilities authority cannot be preserved in this fashion as they are non-duplicable.
+With linear capabilities, authority cannot be preserved in this fashion as they are non-duplicable.
 In order to make a lossless reduction of the range of authority, \trgcm{} provides special hardware support in the form of \texttt{split} and \texttt{splice}.
@@
 Given two capabilities with adjacent ranges of authority and the same permissions and linearity, \texttt{splice} splices them together into one capability.
-The two instructions work in the same way for seal sets.
+The two instructions work in the same way for set-of-seals capabilities.
 We do not provide special support for lossless reduction of capability permissions, but this could probably be achieved with more fine-grained permissions.
@@
   The interpretation of the remainder of the instructions are displayed in Appendix~\ref{app:instr-interpretation}.
-  The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all project information about capabilities.
+  The instructions $\tgetb{}{}$, $\tgete{}{}$, $\tgetlin{}{}$, and $\tgetp{}{}$ all query information about capabilities.
   The $\tgetb{}{}$ and $\tgete{}{}$ instructions, respectively, project the base and end address of the range of authority.
   The linearity of a permission is projected with $\tgetlin{}{}$, and finally the permission is projected with $\tgetp{}{}$.
-  The instructions $\tgetb{}{}$ and $\tgete{}{}$ also work on sets of seals.
+  The instructions $\tgetb{}{}$ and $\tgete{}{}$ also work on set-of-seal capabilities.
   The instruction $\tisptr{}{}$ returns an integer representation of the type of a word (capability or number).
@@
   The latter instruction compares two numbers and writes 1 or 0 to a target register depending on whether one number is less than the other.
-  The instruction $\tsetatob{}$ sets the current address (or seal) of a capability (or set-of-seals) to the base address of the range of authority (or range of seals).
+  The instruction $\tsetatob{}$ sets the current address (or seal) of a memory (or set-of-seals) capability to the base address of the range of authority (or range of seals).
   This instruction makes it easy to work relatively to the base address of a capability.
@@
 
-  Additionally, when the untrusted code returns,  we need to regain access to the capabilities that represent our authority, which we had access to before passing control.
+  Additionally, when the untrusted code returns,  we need to regain access to the capabilities that represent our authority.
   To achieve this, we have to store them in a location that we can access after the return.
-  However, without an encapsulation mechanism, the untrusted code would have access to all this through the return capability because it gives the same authority to the untrusted code as to us after the return.
+  However, without an encapsulation mechanism, the untrusted code would also have the same authority as us, including access to the authority we stored away.
   This is why any reasonable capability machine must have an encapsulation mechanism to allow programs to set up boundaries between security domains.
 
-  In \trgcm{}, we can seal the return capability to establish such a boundary.
+  To establish such a boundary in \trgcm{}, we can seal the return capability.
   Specifically, we can seal the return capability (which points to the instruction to be executed after the return) as a pair, together with a pointer to the stack frame where we have stashed away our capabilities (see further).
@@
 
-  It is worth pointing out that \trgcm{} does not have a direct unsealing instruction to extract a capability from its sealed version when the seal is available, but one can be emulated.
+  It is worth pointing out that \trgcm{} does not have a direct unsealing instruction to extract a capability from its sealed version when the seal is available, but one can be (partially) emulated.
   Say you have a sealed non-executable word $\sealed{\sigma,w}$ as well as a set-of-seals capability $\seal{\sigma_\baddr,\sigma_\eaddr, \sigma}$ that contains $\sigma$, i.e.\ $\sigma \in [\sigma_\baddr,\sigma_\eaddr]$.
@@
   This does not work for sealed executable capabilities because $\txjmp{}{}$ fails if the data capability is executable.
-  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.\footnote{Alternatively, we could remove the current restriction in \texttt{xjmp} that requires the data capability to be non-executable but then teh invoked code would often have to manually perform this check instead.}
+  However, even though we cannot extract the executable capability from $\sealed{\sigma,c}$, we can still invoke it with arbitrary arguments, since we can use the seal to construct an arbitrary data part and $\txjmp{}{}$ to the combination.\footnote{Alternatively, we could remove the current restriction in \texttt{xjmp} that requires the data capability to be non-executable but then the invoked code would often have to manually perform this check instead.}
 
@@
 Rather than defining such a decode function and an encode function, we assume that they are given with certain properties.
-The $\decInstr{} : \Word \fun \Instr$ should decode all capabilities to the $\fail$ instruction, i.e.\ only numbers are decoded to non-fail instructions:
+Most importantly, $\encInstr{} : \Instr \fun \ints$ should be the right inverse of $\decInstr{}$, i.e.
+\[
+\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
+\]
+The $\decInstr{} : \Word \fun \Instr$ should only map numbers to non-$\fail$ instructions, i.e.\ capabilities are mapped to $\fail$:
 \[
   \forall c \in \Caps \ldotp \decInstr{c} = \fail
-\]
-The $\decInstr{}$ should have a right inverse $\encInstr{} : \Instr \fun \ints$, i.e.
-\[
-\forall i \in \Instr \ldotp \decInstr{}(\encInstr{i}) = i
 \]
@@
 \label{subsec:components-linking}
-The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
-To clarify this, we introduce notions of components and programs from which we construct executable configurations.
-A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
-It has exports that can satisfy the imports of other components.
-% Structure
-A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
-The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
-An export associates a word with a symbol.
-%base = library
-A component is either a library component (without a main entry point) or an incomplete program with a main in the form of a pair of sealed capabilities.
-The latter can be seen as a program that still needs to be linked with libraries.
-% Linking .
-Components are combined into new components by linking them together, as long as only one has a main function.
-Two components can be linked when their memories, seals, and linear addresses are disjoint.
-They are combined by taking the union of each of their constituents.
-For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
-The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.
-
-% Programs
-We can now define the notion of a program as well as a context.
-\begin{definition}[Programs and Contexts]
-  \label{def:program-and-context}
-  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
-  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
-\end{definition}
-% Getting an executable from a component - postpone this to later, but mention:
-% write xor exec
-How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
-% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
-%Further, we will only consider a specific class of well-formed components, but we will get back to this.
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.
-\begin{figure}[tb]
+
+\begin{figure}
   \centering
@@
 
+The executable configuration describes the machine state, but it does not make it clear what components run on the machine and how they interact with each other.
+To clarify this, we introduce notions of components and programs from which we construct executable configurations.
+A component (defined in Figure~\ref{fig:target-component-and-linking}) is basically a program with entry points in the form of imports that need to be linked.
+It has exports that can satisfy the imports of other components.
+% Structure
+A base component $\comp_0$ consists of a code memory fragment, a data memory fragment, a list of imported symbols, a list of exported symbols, two lists specifying the available seals (see Section~\ref{sec:form-secur-with}), and a set of all the linear addresses (addresses governed by a linear capability).
+The import list specifies where in memory imports should be placed, and imports are matched to exports via their symbols.
+An export associates a word with a symbol.
+%base = library
+A component is either a library component (without a main entry point) or an incomplete program with a main entry point in the form of a pair of sealed capabilities.
+The latter can be seen as a program that still needs to be linked with libraries.
+% Linking .
+Components are combined into new components by linking them together, as long as only one has a main function.
+Two components can be linked when their memories, seals, and linear addresses are disjoint.
+They are combined by taking the union of each of their constituents.
+For every import that is satisfied by an export of the other component, the data memory is updated to have the exported word on the imported address.
+The satisfied imports are removed from the import list in the resulting linked component and the exports are obtained by combining the components' exports.
+
+% Programs
+We can now define the notion of a program as well as a context.
+\begin{definition}[Programs and Contexts]
+  \label{def:program-and-context}
+  A \emph{program} is a component $(\comp_0,c_{\mathrm{main},c}, c_{\mathrm{main},d})$ with an empty import list.
+  A \emph{context} for a component $\comp$ is a component $\comp'$ such that $\comp \bowtie \comp'$ is a program.
+\end{definition}
+% Getting an executable from a component - postpone this to later, but mention:
+% write xor exec
+How a program is initialised to create an executable configuration is discussed in Section~\ref{sec:form-secur-with}.
+% We will, however, at this point mention that all programs will be instrumented as Write-XOR-Execute,{Do we have a reference for this?} i.e.\ there will only be read-execute capabilities for the code and read-write capabilities for the data.
+%Further, we will only consider a specific class of well-formed components, but we will get back to this.
+Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.
+
 %\FloatBarrier
@@
 The problem is essentially that the stack and return pointers that a callee receives from a caller remain in effect outside their intended lifetime: either when the callee has already returned or when they have themselves invoked other code.
-Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the receiver cannot have kept a copy of it as it is non-duplicable.} that can be used to prevent this from happening.
+Linear capabilities offer a form of revocation\footnote{Revocation in the sense that if we hand out a linear capability and later get it back, then the adversary cannot have kept a copy of it as it is non-duplicable.
+In other words, the adversary's access has effectively been revoked in this situation.} that can be used to prevent this from happening.
 
@@
 Because the stack capability is linear, the caller knows that the capability for their local stack frame cannot have an alias.
-This means that an adversary would need the stack capability produced by the caller in order to access their local data.
+This means that an adversary cannot access the caller's local data because the caller has a linear capability for it and there cannot exist an alias.
 The caller gives this capability to the adversary only in a sealed form, rendering it opaque and unusable.
@@
 A caller pushes a stack frame to the stack on call and a callee pops a stack frame from the stack upon return.
-However without any enforcement, there is nothing to prevent a callee from returning from an arbitrary call on the call stack.
+However without any enforcement, there is nothing to prevent a callee from popping more from the stack than they should and returning on an arbitrary call on the call stack.
 This is exactly what the adversary does in Figure~\ref{fig:ret-ptr-abuse} when they skip two stack frames.
@@
 In \stktokens{}, the base address of the stack capability is fixed as a compile-time constant (Note: the stack grows downwards, so the base address of the stack capability is the top-most address of the stack).
-The caller verifies the validity of the return token by checking whether the base address of a returned token corresponds to this fixed base address, which was the base address for the return token they gave to the callee.
-In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.
+The caller verifies the validity of the return token by checking whether the base address of a returned token corresponds to this fixed base address.
+In the scenario we just sketched, the base address check would fail in Figure~\ref{fig:stack-base-abuse-d}, thus alerting the caller to the attempt to break WBCF.
 
@@
 % Don't be stupid
-Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
+Return seals should also never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller and thus break LSE.
 This goes for direct leaks (leaving a seal in a register or writing it to adversarial memory), as well as indirect leaks (leaking a different capability that can be used for reading, either directly or indirectly, a return seal from memory).
@@
 \item \label{item:return-data} Create token and data return capability on call: split the stack capability in two to get a stack capability for your local stack frame and a stack capability for the unused part of the stack.
-The former is sealed and used for the data part of the return pair.
+The former is sealed and used as the data part of the return pair.
 The latter is passed to the callee as the stack pointer.
 \item \label{item:return-code} Create code return capability on call: Seal the old program counter capability.
-\item Reasonable use of seals: Return seals are only used to seal old program counter capabilities, every return seal is only used for one call site, and they are not leaked.
+\item Reasonable use of seals: Return seals are only used to seal old program counter capabilities, every return seal is only used for one call site, and they are not leaked (directly or indirectly).
 \end{enumerate}
@@
 
-The main difference between the two semantics is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built-in to the abstract machine.
+The main difference between the two semantics is that the new one satisfies LSE and WBCF by construction: the abstract machine comes with a built-in stack, inactive stack frames are unaddressable and well-bracketed control flow is built into the abstract machine.
 Important run-time values like return capabilities and stack pointers are represented by special syntactic tokens that interact with the abstract machine's stack, but during execution, there remains a close, structural correspondence to the actual regular capabilities that they represent.
@@
 Every stack frame contains encapsulated stack memory as well as the program point that execution is supposed to return to.
-The free stack memory is the active part of the stack that has not been claimed by a call and thus can be used at this point of time.
+The free stack memory is the active part of the stack that has not been encapsulated by a previous call and thus can be used by the currently executing code.
 % Stack pointer
@@
 % Return pointers
-The final syntactic constructs added by \srccm{} are the code and data return pointers.
-The data return pointer corresponds to some stack pointer (which in turn corresponds to a linear capability), and the code return pointer corresponds to some capability with read-execute permission.
-Syntactically, the return capabilities contain just enough information to reconstruct what they correspond to on the underlying machine.
-On \srccm{}, return pointers are generated by calls from the capabilities they correspond to on \trgcm{}, and they are turned back to the capabilities they correspond to upon return.
+The final syntactic constructs added by \srccm{} are code and data return pointers.
+The data return pointer corresponds to a stack capability, and the code return pointer corresponds to a executable capability for the corresponding return point.
+Syntactically, the return pointers contain just enough information to reconstruct what they correspond to on the underlying machine.
+During \srccm{} function calls, return pointers are generated from the stack and program counter capabilities, and they are turned back to those capabilities upon return.
 % Even though the return pointers both correspond to capabilities with read permission, neither can be used for reading. 
@@
 
-The opaque nature of the return pointers is reflected in the interpretation of the instructions common to both \trgcm{} and \srccm{} as \srccm{} does not add special interpretation for them in non-\texttt{xjmp} instructions.
+The opaque nature of the return pointers is reflected in the interpretation of instructions as \srccm{} does not add special interpretation for them in non-\texttt{xjmp} instructions.
 Stack pointers, on the other hand, need to behave just like capabilities, so \srccm{} adds new cases for them in the semantics, e.g.\ \texttt{cca} can now also change the current address of a stack pointer as displayed in Figure~\ref{fig:source-op-sem}.
 Similarly, \texttt{load} and \texttt{store} work on the free part of the stack when provided with a stack pointer.
-A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view the overlay semantics has on the underlying machine.
-In other words, there should only be stack pointers for the stack memory.
+A store attempted with a stack capability that points to an address outside the free stack results in the $\failed$ configuration because that action is inconsistent with the view of the overlay semantics on the underlying machine.
+In \srccm{} executions, there should only be stack pointers for the stack memory.
 
@@
 To formalize this distinction, \srccm{} has a set of trusted addresses $\ta$.
-Only instructions at these addresses can be interpreted as the \srccm{} native call and push frames to the call stack which guarantees LSE and WBCF.
-The constant $\ta$ is a parameter of the \srccm{} step relation.
-Similarly, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
-
-Apart from the step relation of \trgcm{}, \srccm{} has one overlay step that takes precedence over the others.
-This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than one.
-The sequence of instructions have to correspond to a call, i.e.\ the
-instructions in Figure~\ref{fig:call-code} ({\footnotesize  $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$'th instruction in the figure and $\calllen$ is always $26$, i.e.\ the number of instructions).
+This $\ta$ is a constant parameter of the \srccm{} step relation.
+Only instructions at these addresses will be interpreted as native \srccm{} calls and push frames to the call stack.
+As such, WBCF and LSE for calls from these addresses.
+In addition to $\ta{}$, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.
+
+In addition to the (modified) reduction rules of \trgcm{}, \srccm{} has an overlay step that takes precedence over the others.
+This step is shown in Figure~\ref{fig:source-op-sem}, and it is different from the others in the sense that it interprets a sequence of instructions rather than a single one.
+The sequence of instructions have to correspond to a call, i.e.\ the instructions in Figure~\ref{fig:call-code} ({\footnotesize  $\scall[i]{\offpc,\offsigma}{r_1}{r_2}$} corresponds to the $i$th instruction in the figure and $\calllen$ is always $26$, i.e.\ the number of instructions).
 Calls are only executed when the well-behaved component executes, so the addresses where the call resides must be in $\ta$, and the executing capability must have the authority to execute the call.
@@
 $\scall{}{}{}$ also constructs a return capability $c_{\var{opc}}$ and its address $\var{opc}$, pointing after the call instructions.
-The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of sealed return capabilities.
-The return capabilities are sealed with the seal designated for the call.
-
-The return capabilities, $\retptrc$ and $\retptrd$ are sealed and can only be used using the \texttt{xjmp} instruction, to perform a return.
+The local stack frame and return address are pushed onto the stack, and the local stack capability and return capability are converted into a pair of return capabilities and sealed with the seal designated for the call.
+
+Since the return capabilities $\retptrc$ and $\retptrd$ are sealed, they can only be used using the \texttt{xjmp} instruction, to perform a return.
 When this happens, the topmost call stack frame $(\opc,\ms_\var{local})$ is popped from the call stack.
@@
 A tail call is a call from a caller that is done executing, and thus doesn't need to be returned to or preserve local state.
-This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in call.
-To perform a tail call, the caller simply transfers control to the callee using \texttt{xjmp}.
+This means that a tail call should not reserve a slot in the return order by pushing a stack frame on the call stack, i.e.\ it should not use the built-in \srccm{} call.
+Instead, to perform a tail call, the caller simply transfers control to the callee using \texttt{xjmp}.
 The tail-callee should return to the caller's caller, so the caller leaves the return pair they received for the callee to use.
@@
 
+  Most of the definitions in this chapter will be parameterized by a tuple $(\ta,\stkb,\gsigrets,\gsigcloss)$ that we will sometimes denote with the meta-variable $\gc$.
+  This tuple collects four \emph{global constants}.
+  In addition to $\ta$, the set of trusted addresses, and $\stkb$, the fixed base address of the stack memory, which we have already encountered in Section~\ref{subsec:overlay-semantics}, the tuple contains two additional constants: $\gsigrets$ and $\gsigcloss$: respectively the set of all return and closure seals in the system.
+
   \subsubsection{Reasonable words}
@@
   \label{def:reasonable-word}
-  Take a set of trusted addresses $\ta$ and sets of return and closure seals $\gsigrets$ and $\sigcloss$.
+  Take a set of trusted addresses $\ta$ and sets of return and closure seals $\gsigrets$ and $\gsigcloss$.
 We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
@@
 Finally, we define that stack or memory capabilities with read permission are reasonable if the memory they give access to contains only reasonable words.
-To accomodate this, reasonability of words is defined relative to the memory and stack.
+To accommodate this, reasonability of words is defined relative to the memory and stack.
 The definition is cyclic, but the step-index ensures that it is well-founded.
@@
 \label{sec:related-work}
-In this section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation or the linear capabilities we use to do it.
+In this final section, we discuss related work on securely enforcing control flow correctness and/or local state encapsulation or the linear capabilities we use to do it.
 We do not repeat the work we discussed in Section~\ref{sec:introduction}.
@@
 If we see reasonable behavior as defined behavior, then our full abstraction result can be seen as an application of this same idea.
-
 In follow-up work, \citet{Abate:2018:GCG:3243734.3243745} extend this approach to components with  undefined behavior by proving secure compilation guarantees for the execution until the undefined behavior happens.
+
 Another interesting relation between our work and that of \citet{Abate:2018:GCG:3243734.3243745}, is that they consider dynamic compromise scenarios, where some components start out as trusted until they are compromised and are treated as adversarial after that.
@@
 This can be seen as a step-bounded version of the idea explained in \citep{patrignani_modular_2016}.
-It is an interesting observation that modular secure compilation, combined with step-bounded reasoning is sufficient to cover dynamic compromise scenarios without the need to build in dynamic compromise scenarios into the definition of secure compilation like \citet{Abate:2018:GCG:3243734.3243745}.
+If confirmed by a more careful analysis, it would be an interesting observation that modular secure compilation, combined with step-bounded reasoning is sufficient to cover dynamic compromise scenarios without the need to build in dynamic compromise scenarios into the definition of secure compilation like \citet{Abate:2018:GCG:3243734.3243745}.
 
@@
 \citet{tsampas_2019} allow (almost) normal references that can be stored in multiple places at the same time.
-This means that their approach is more like C than \stktokens{}.
+This means that their approach is closer to C than \stktokens{}.
 As explained in Section~\ref{sec:introduction}, these approaches have the downside that they require stack clearing (including unused parts) on boundary crossings.

### 2020-12-27 popl-2019-journal-v/discussion.tex
Fix another reviewer remark.

@@
   When it comes to fully-abstract compilation, then the answer is that all properties in the
-  high-level language should be enforced (or at least the equivalences they imply), so the real question is what high-level language we would like.
+  high-level language should be enforced (or at least the equivalences they imply\footnote{See Section~\ref{sec:related-work} for a discussion of secure compilation properties that require preservation of more general language properties than equivalences.}) must be preserved in the target language, either by translating them to similar abstractions in the target language or by using the available target language features to enforce the source language abstraction.
+  In our case, \srccm{} semantics offers a native stack with LSE and WBCF, but this abstraction does not natively exist in \trgcm{}.
+  In order to enforce the abstraction on \trgcm{}, \stktokens{} can be used.
+  Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these abstraction properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
+), so the real question is what high-level language we would like.
   \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \texttt{longjmp}, then we need to come up with a different enforcement scheme.

### 2020-12-27 popl-2019-journal-v/paper.tex
Fix another reviewer remark.

@@
 %Further, we will only consider a specific class of well-formed components, but we will get back to this.
-Some simplifications have been made in this presentation of \trgcm{}. See \citet{technical_report_popl} for details.
+Some simplifications have been made in this presentation of \trgcm{}.
+These simplifications concern minor details, for example, we've omitted the fact that \texttt{move}s to the \texttt{pc} register are disallowed.
+Full details can be checked in \citet{technical_report_popl}.
 
@@
 Generally speaking, behavioral equivalences are preserved and reflected by fully-abstract compilers.
-This means that any property the source language has must somehow be there after compilation, whether or not it is a property of the target language.
-If the source language has a property that the target language doesn't have, then a compiled source program must use the available target language features to emulate the source language property in a way that exactly matches behaviorally.
-In our case, LSE and WBCF are built into the semantics of \srccm{}, but they are not properties of \trgcm{}.
-In order to enforce these properties, components on \trgcm{} use \stktokens{}.
-Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
+This means that source language abstractions (or at least the equivalences they imply\footnote{See Section~\ref{sec:related-work} for a discussion of secure compilation properties that require preservation of more general language properties than equivalences.}) must be preserved in the target language, either by translating them to similar abstractions in the target language or by using the available target language features to enforce the source language abstraction.
+In our case, \srccm{} semantics offers a native stack with LSE and WBCF, but this abstraction does not natively exist in \trgcm{}.
+In order to enforce the abstraction on \trgcm{}, \stktokens{} can be used.
+Theorem~\ref{thm:full-abstraction} proves that \stktokens{} enforces these abstraction properties in a way that behaviorally matches \srccm{} which means that it enforces LSE and WBCF.
 

### 2020-12-28 popl-2019-journal-v/paper.tex
Add a paragraph on Laird's full abstraction result.

@@
 
+\begin{jversion}
+  The use of return pointers or capabilities in low-level languages is in many ways similar to lambda calculus terms in continuation passing style.
+  When applying a CPS-translation to lambda terms, well-bracketedness can also be violated if continuations are invoked more than once.
+  Enforcing well-bracketedness for CPS terms means enforcing that continuations are only used a single time and only in the right invocation, i.e.\ a form of linear (or more accurately, affine) usage of continuations must be enforced.
+  One way to enforce this is by assigning well-chosen types to CPS-translated terms and researchers have proven effectiveness for two ways of doing this, in the form of a full-abstractness result for the CPS-translation.
+  In both cases, correct usage of continuations is enforced by assigning well-chosen types to CPS-translated terms.
+  In the case of \cite{laird_game_2005}, an affine type system is used to do this, while \citet{ahmed_equivalence-preserving_2011} use a form of answer-type polymorphism instead.
+  These results are related to ours because they enforce well-bracketedness using a form of linearity of continuations or return capabilities, but there are also quite a few differences: the static nature of the enforcement, the low-level vs.\ high-level nature of the languages, the fact that LSE comes naturally in such a setting while we also have to enforce it.
+Additionally, \citeauthor{ahmed_equivalence-preserving_2011} consider total languages and use a back-translation in two steps using an intermediate multi-language and \citeauthor{laird_game_2005} proves full abstraction using a fully abstract denotational semantics for source and target languages.
+\end{jversion}
+
 The fact that our full abstraction result only applies to reasonable components (see Section~\ref{sec:form-secur-with}) makes it related to full abstraction results for unsafe languages.

### 2020-12-29 popl-2019-journal-v/proving-full-abstraction.tex
Full pass over paper, fix another reviewer remark.
Fix some layout issues and typos.

@@
 
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+  \begin{tikzpicture}[scale=.45, every node={scale=.5}]
     \begin{scope}
@@
 % $\lrfree[]$ also makes sure that the base address of the stack is actually part of the free stack.
+
+Note that the judgement $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ (and other definitions in this section) is parameterized by $\gc = (\ta,\stkb,\gsigrets,\gsigcloss)$, the global constants that we previously encountered in Section~\ref{sec:reasonable-components} when discussing reasonable components.
 
@@
   \centering
-  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
+  \begin{tikzpicture}[scale=.4, every node={scale=.5}]
 
@@
 It takes related memories and register files (ignoring the $\rdata$ register) and combines them into two configurations.
-Each of the configurations is plugged with a code capability and a data capability just like the $\txjmp{}{}$ instruction would do it and requires the resulting configurations to be in the $\lroname$ relation.
+Each of the configurations is plugged with a code and data capability just like the $\txjmp{}{}$ instruction would and requires the resulting configurations to be in the $\lroname$ relation.
 The $\lrename_\mathrm{xjmp}$ relation is defined in \figurename~\ref{fig:expr-rels}.
@@
     \begin{array}{l}
-      \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
-      \quad \forall W_\lrrs , W_\lrm \ldotp \\
+      \forall n' \leq n, W_\lrrs , W_\lrm, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk} \ldotp\\
       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs) \wedge\\
@@
 
-\caption{The expression relation relates capabilities capabilities that can safely be used for execution.
-The xjmp expression relation can be used to relate capabilities that are safe as sealed capabilties.}
+\caption{The expression relation relates capabilities that can safely be used for execution. The xjmp expression relation relates capabilities that are safe as sealed capabilties.}
 \label{fig:expr-rels}

### 2021-02-21 popl-2019-journal-v/discussion.tex
Fix final reviewer comments.

--- 2020-12-27 popl-2019-journal-v/discussion.tex
@@
   A compiler is secure when it enforces the properties of high-level languages which begs the question what properties we should enforce.
-  When it comes to fully-abstract compilation, then the answer is that all properties in the
-  high-level language should be enforced (or at least the equivalences they imply), so the real question is what high-level language we would like.
+  When it comes to fully-abstract compilation, then the answer is that all properties in the high-level language (or at least the equivalences they imply\footnote{See Section~\ref{sec:related-work} for a discussion of secure compilation properties that require preservation of more general language properties than equivalences.}) must be preserved in the target language, either by translating them to similar abstractions in the target language or by using the available target language features to enforce the source language abstraction.
+  In our case, \srccm{} semantics offers a native stack with LSE and WBCF, but this abstraction does not natively exist in \trgcm{}, so \stktokens{} can be used to enforce it.
+
+  Perhaps a more subjective question is what kind of high-level language we would like.
   \stktokens{} ensures a standard call-return control-flow, but if we want a different kind of control-flow, for instance call/cc or C's \texttt{longjmp}, then we need to come up with a different enforcement scheme.

### 2021-02-21 popl-2019-journal-v/paper.tex
Fix final reviewer comments.

@@
 % Cap machine inspired by CHERI, but with linear capabilities
-In this section, we introduce a simple but representative capability machine with linear capabilities, that we call \trgcm{} (Linear Capability Machine).
+In this section, we introduce a simple but representative capability machine with linear capabilities, which we call \trgcm{} (Linear Capability Machine).
 \trgcm{} is mainly inspired by CHERI~\citep{watson_cheri_2015} with linear capabilities as the main addition.
@@
 Only instructions at these addresses will be interpreted as native \srccm{} calls and push frames to the call stack.
-As such, WBCF and LSE for calls from these addresses.
+As such, WBCF and LSE will be guaranteed only for calls from these addresses.
 In addition to $\ta{}$, \stktokens{} assumes a fixed base address of the stack memory, that is also passed around as such a parameter, for use in the native semantics of calls.

### 2021-02-21 popl-2019-journal-v/proving-full-abstraction.tex
Fix final reviewer comments.

@@
 \]
+For ease of explanation, we will gradually introduce the definition of these three components, starting with simplified versions and gradually introducing different complications before obtaining the intended definition of worlds by the end of Section~\ref{subsec:worlds}.
+That version will only be modified a bit further in Section~\ref{sec:rec-dom-eq}, purely for technically justifying the recursive definition of the worlds.
+The final definitions can be found in Theorem~\ref{thm:rec-dom-eq-sol}.
 
@@
 We use the world to do this by having two kinds of regions: shared and spatial.
-Memory governed by shared and spatial regions can only be references by non-linear and linear capabilities respectively.
+Memory governed by shared and spatial regions can only be referenced by non-linear and linear capabilities respectively.
 
@@
 However, spatial regions are not enough.
-A world which contains a spatial region, represents the assumption of exclusive ownership of the corresponding memory.
+A world that contains a spatial region, represents the assumption of exclusive ownership of the corresponding memory.
 In other words, a linear capability for this memory will be valid only in such a world.
@@
 This setup will allow ownership of a memory fragment to reside with linear capabilities which are not themselves stored in the memory fragment but elsewhere in memory or in a register.
-In such a setup, the region representing the memory fragment will exist only as $\spatial$ in the world where the memory fragment's contents is valid, but it will exist as $\spatialo$ in the world where the linear capability owning the fragment is valid.
+In such a setup, the region representing the memory fragment will exist only as $\spatial$ in the world where the memory fragment's contents are valid, but it will exist as $\spatialo$ in the world where the linear capability owning the fragment is valid.
 
@@
 For example, two related capabilities remain related when an extra invariant is added on unrelated memory, or when a stack frame that it does not reference is dropped.
-In Kripke logical relations, such changes that do not invalidate the relatedness of values, are modelled by the future world relation $\future$.
+In Kripke logical relations, such changes that do not invalidate the relatedness of values are modelled by the future world relation $\future$.
 The future world relation can also be thought of as the model of allowed changes in memory over time.
@@
 As such, we require the world-indexed memory invariants to be monotone in the world:
+\[
+  \Regions = \left\{
+  \begin{array}{l}
+    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
+    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg})
+  \end{array} \right.
+\]
+We make a similar change to $\Regionh$. The seal invariants must be monotone as well.
+
+Kripke future world relations usually allow extending worlds with extra assumptions, or take steps in protocols that the system was designed to follow.
+However, in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on this assumption any more.
+Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory and drop or replace the previous assumption.
+We mark dropped regions as revoked in the world following \citet{ahmed_2004} and \citet{thamsborg_2011} which for all intents and purposes corresponds to actually dropping the region.
+
+We add a $\revoked$ tag to the spatial regions $\Regions$:
 \[
@@
 \]
-We make a similar change to $\Regionh$. The seal invariants must be monotone as well.
-
-Kripke future world relations usually allow extending worlds with extra assumptions, or take steps in protocols that the system was designed to follow.
-However, in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on this assumption any more.
-Specifically, if we have the only linear capability for a piece of memory, then we can be sure that there are no other capabilities for the same memory which makes it safe to repurpose the memory and drop or replace the previous assumption.
-We mark dropped regions as revoked in the world following \citet{ahmed_2004} and \citet{thamsborg_2011} which for all intents and purposes corresponds to actually dropping the region.
-
-We add a $\revoked$ tag to the spatial regions $\Regions$:
-\[
-  \Regions = \left\{
-  \begin{array}{l}
-    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\
-    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\
-    \{\revoked \}
-  \end{array} \right.
-\]
-The $\spatialo$-region may be depended on by a linear capability \emph{may} so we cannot allow it to be revoked.
+The $\spatialo$-region may be depended on by a linear capability so we cannot allow it to be revoked.
 On the other hand, no capability can depend on a $\spatial$-region, so it can be safely revoked.
@@
 
-\caption{Example what happens to worlds when a linear capability $c_{\var{lin}}$ is repurposed.
+\caption{An example of what happens to worlds when a linear capability $c_{\var{lin}}$ is repurposed.
   The linear capability is originally valid with respect to $W_1$.
@@
 \begin{multline*}
-    \npair{(\overline{\sigma},\src{ms},\ms_T)} \in \lrheap(\pwheap)(W') =\\
+    (\overline{\sigma},\src{ms},\ms_T) \in \lrheap(\pwheap)(W') =\\
     \left\{
@@
         \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
-        \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
-        \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+        \qquad \forall r \in \dom(\activeReg{\pwheap}) \ldotp \\
+        \qquad \quad R_\ms(r) \in  \pwheap(r).H \; R_W(r) \wedge\\
         \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
@@
   \begin{multline*}
-    \memSatFStack{\src{\ms_\stk},\ms_T}{W} \text{ iff } \\
+    \memSatFStackNoN{\src{\ms_\stk},\ms_T}{W} \text{ iff } \\
     \left\{
@@
         \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge\\
-        \qquad \forall r \in \dom(\activeReg{W_\var{stack}}),n' < n \ldotp \\
-        \qquad \quad \npair[n']{R_\ms(r)} \in  W_\var{stack}(r).H \; \xi^{-1}(R_W(r))
+        \qquad \forall r \in \dom(\activeReg{W_\var{stack}}) \ldotp \\
+        \qquad \quad R_\ms(r) \in  W_\var{stack}(r).H \; R_W(r)
       \end{array}
@@
   \begin{multline*}
-    \memSatStack{\src{\stk},\ms_T}{W} \text{ iff } \\
+    \memSatStackNoN{\src{\stk},\ms_T}{W} \text{ iff } \\
     \left\{
@@
           \quad\qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
-          \quad\qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n\ldotp \\
-          \quad\qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
+          \quad\qquad \forall r \in \dom(\activeReg{W_\var{stack}})\ldotp \\
+          \quad\qquad \quad (\pi_1(R_\ms(r)),\pi_3(R_\ms(r)) \in W_\var{stack}(r).H \; R_W(r) \wedge\\
           \quad\qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
@@
   \begin{multline*}
-      \memSat{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff }\\
+      \memSatNoN{\src{\ms_S},\src{\ms_\stk},\src{\stk},\ms_T}{W} \text{ iff }\\
       \left\{
@@
           \qquad \baddr_\stk -1,\eaddr_\stk + 1 \in \dom(\ms_{T,f}) \wedge \\
-          \qquad \memSatStack{\src{\stk},\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
-          \qquad \memSatFStack{\src{\ms_\stk},\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
-          \qquad \npair{(\overline{\sigma},\src{\ms_S'},\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
+          \qquad \memSatStackNoN{\src{\stk},\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
+          \qquad \memSatFStackNoN{\src{\ms_\stk},\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
+          \qquad (\overline{\sigma},\src{\ms_S'},\ms_\var{T,heap}) \in \lrheap(\pwheap)(W_{\var{heap}})
         \end{array}
@@
 \]
+
+For ease of presentation, we have omitted step indices and the isomorphisms $\xi$ from the definitions in Section~\ref{subsubsec:mem-sat}.
 
@@
 
-Sometimes we need to know that the contents of a memory segment stays the same.
+Sometimes we need to know that the contents of a memory segment stay the same.
 For instance, the contents of encapsulated stack frames do not change which we need to be able to rely on.
@@
 \]
-The code region is $\pure$ because it needs to contain a seal invatiant and because we assume that code pointers are normal capabilities.
+The code region is $\pure$ because it needs to contain a seal invariant and because we assume that code pointers are normal capabilities.
 

### 2021-02-24 popl-2019-journal-v/paper.tex
Add some very recent additional references.

@@
 
-Local capabilities can be used to ensure well-bracketed control-flow and local-state encapsulation as demonstrated by \citet{skorstengaard_reasoning_2017}.
+Local capabilities can be used to ensure well-bracketed control-flow and local-state encapsulation as demonstrated by \citet{skorstengaard_reasoning_2017,skorstengaard_reasoning_2019}.
 Recently, \citet{tsampas_2019} demonstrated that an extension of local capabilities with multiple linearly ordered colours can be used to enforce the life time of stack references.
@@
 As explained in Section~\ref{sec:introduction}, these approaches have the downside that they require stack clearing (including unused parts) on boundary crossings.
+Very recently, \citet{georges_efficient_2021} have proposed a version of \citeauthor{skorstengaard_reasoning_2017}'s calling convention which succesfully avoids the need for stack clearing by relying on a newly proposed form of capabilities called uninitialized capabilities and they have proven soundness of the approach using Coq and the Iris program logic \citep{jung_iris_2018}.
 
